\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{array}
\usepackage{caption}
\usepackage{changepage}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{framed}
\usepackage{mdframed}

\pagestyle{fancy}
\fancyhf{}
\rhead{Fate's Edge}
\lhead{Adventure Module Template}
\rfoot{\thepage}

\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

\newcommand{\player}[1]{\textbf{#1:}}
\newcommand{\gm}[1]{\textit{GM: #1}}
\newcommand{\roll}[2]{\texttt{Roll: #1d10 = [#2]}}
\newcommand{\success}[1]{\textbf{Successes: #1}}
\newcommand{\cp}[1]{\textbf{SB: #1}}
\newcommand{\outcome}[1]{\textbf{Outcome: #1}}
\newcommand{\clocksegment}{\textbullet}

\definecolor{clockcolor}{rgb}{0.2,0.4,0.8}
\definecolor{modulebg}{rgb}{0.95,0.95,0.95}

\begin{document}

\begin{center}
{\LARGE Fate's Edge Adventure Module Template}\\
\vspace{0.5em}
{\large Creating Cohesive Scenarios Using Deck Inspiration}
\end{center}

\section*{Adventure Module Structure Overview}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{Core Components of Every Module:}
\begin{enumerate}[leftmargin=*]
\item \textbf{Theme and Tone}: What feeling should this evoke?
\item \textbf{Deck-Based Seed}: Using card draws to inspire cohesive elements
\item \textbf{Narrative Hook}: Why do the PCs care?
\item \textbf{Key NPCs}: Who populates this story?
\item \textbf{Scene Clocks}: What ongoing pressures drive tension?
\item \textbf{Resolution Paths}: How can PCs succeed (or fail) meaningfully?
\item \textbf{Campaign Integration}: How does this fit into larger arcs?
\end{enumerate}
\end{mdframed}

\section*{Template: "The Singing Wastes" - A Black Banners Adventure}

\subsection*{1. Theme and Tone}

\begin{itemize}[leftmargin=*]
\item \textbf{Theme}: War's aftermath, haunted battlefields, the cost of violence
\item \textbf{Tone}: Melancholy, mysterious, with moments of terrible beauty
\item \textbf{Genre Elements}: Military fantasy, supernatural mystery, post-conflict drama
\item \textbf{Mood Words}: Haunting, desolate, echoing, frozen, restless
\end{itemize}

\subsection*{2. Deck-Based Seed Generation}

\player{GM} Using the Black Banners generator to create our scenario:

\gm{Drawing until all four suits appear:}

\begin{description}[leftmargin=*]
\item[Spade (Q - The Bone Fields)] A place where the Ykrul clans come to negotiate terms with the dead
\item[Heart (J - The Black Colonel)] Legendary commander who hasn't been seen in the field for months  
\item[Club (K - High Chief plays the companies against each other)] Winner takes nothing
\item[Diamond (Q - Clan-Mother's war-braid)] Her warriors will fight beside you this battle
\end{description}

\gm{Highest rank is Q (Queen), so we have an 8-segment primary clock. The Ace rule applies: weapons that remember, ice that holds the dead, loyalty that needs no banner.}

\subsection*{3. Narrative Hook}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{The Premise}: The PCs are hired by a Vilikari noble to investigate why the Bone Fields - a traditional neutral ground where war-bands come to negotiate with the dead - have become dangerous. Mercenary companies are disappearing, and the Black Colonel (a legendary figure thought dead) has been sighted leading ghostly troops.

\textbf{PC Motivations}:
\begin{itemize}[leftmargin=*]
\item \textbf{Mercenary}: Former soldier seeking answers about missing comrades
\item \textbf{Noble}: Family honor at stake, political implications
\item \textbf{Scholar}: Ancient battlefields hold historical/magical significance
\item \textbf{Outsider}: Curious about the strange customs of the borderlands
\end{itemize}
\end{mdframed}

\subsection*{4. Key NPCs}

\begin{description}[leftmargin=*]
\item[The Black Colonel (Antagonist/Enigma)] Once a brilliant tactician, now a wraith bound to the battlefield. Not evil, but unable to stop fighting the war he never finished. Carries a banner that makes the dead fight. 
\begin{itemize}
\item \textbf{Motivation}: Finish the battle that cost him everything
\item \textbf{Method}: Uses the Clan-Mother's war-braid to bind spirits to his cause
\item \textbf{Weakness}: Cannot cross running water; obsessed with honor
\end{itemize}

\item[Clan-Mother Yara (Allies/Obstacle)] A Ykrul leader who initially hired the Black Colonel but now regrets it. She's pragmatic and honorable, but her war-braid has been corrupted.
\begin{itemize}
\item \textbf{Motivation}: Save her people from the Colonel's influence
\item \textbf{Resources}: Knowledge of death-rites, loyal warriors
\item \textbf{Complication}: Her own ancestors now fight against her
\end{itemize}

\item[Captain Thorne (Red Herring/Ally)] Leader of a missing Free Company. Found alive but changed - he speaks of beautiful music that called his men away.
\begin{itemize}
\item \textbf{Motivation}: Wants to return to his men, doesn't understand he's part of the haunting
\item \textbf{Knowledge}: Saw the Black Colonel's banner, heard the singing
\item \textbf{Role}: Guide to the deeper mystery, potential sacrifice
\end{itemize}
\end{description}

\subsection*{5. Scene Clocks}

\begin{center}
\textbf{Singing Wastes Corruption Clock} (8 segments)\\
\fbox{\clocksegment\clocksegment\clocksegment\clocksegment\clocksegment\clocksegment\clocksegment\clocksegment} 0/8\\
\textit{How far has the haunting spread beyond the Bone Fields}
\end{center}

\textbf{Advancement Triggers}:
\begin{itemize}[leftmargin=*]
\item 1-2 SB: Local spirits become restless
\item 3-4 SB: Living begin hearing the songs
\item 5-6 SB: Dead start appearing in nearby settlements
\item 7-8 SB: River crossing becomes impassable, region quarantined
\end{itemize}

\begin{center}
\textbf{Colonel's Army Clock} (6 segments)\\
\fbox{\clocksegment\clocksegment\clocksegment\clocksegment\clocksegment\clocksegment} 0/6\\
\textit{Strength of the ghostly force under the Black Colonel's command}
\end{center}

\textbf{Advancement Triggers}:
\begin{itemize}[leftmargin=*]
\item Each missing company: +1 segment
\item Successful death-rite by the Colonel: +1 segment
\item PCs fail to disrupt a haunting: +1 segment
\item Clan-Mother's resistance weakens: +2 segments
\end{itemize}

\begin{center}
\textbf{Clan-Mother's Resolve Clock} (4 segments)\\
\fbox{\clocksegment\clocksegment\clocksegment\clocksegment} 0/4\\
\textit{Her willingness to sacrifice to end the haunting}
\end{center}

\textbf{Advancement Triggers}:
\begin{itemize}[leftmargin=*]
\item Seeing ancestors among the enemy: +1 segment
\item Pressure from Vilikari noble: +1 segment
\item Personal losses accumulate: +1 segment
\item PCs show weakness/mistakes: +1 segment
\end{itemize}

\subsection*{6. Adventure Flow}

\subsubsection*{Act I: Arrival and Discovery (Sessions 1-2)}

\textbf{Opening Scene}: The PCs arrive at the Bone Fields to find them unnaturally quiet. The traditional markers and burial mounds are intact, but no war-bands are present.

\textbf{Key Scenes}:
\begin{enumerate}[leftmargin=*]
\item Investigation of abandoned campsites (Wits + Survival/Lore)
\item First encounter with the haunting - ghostly music, cold spots
\item Meeting Captain Thorne, who warns them away
\item Discovery of the Black Colonel's banner in a makeshift shrine
\end{enumerate}

\textbf{Clock Advancement}: 
\begin{itemize}[leftmargin=*]
\item Singing Wastes Corruption: +2 (first haunting encounters)
\item Colonel's Army: +1 (banner discovered)
\end{itemize}

\subsubsection*{Act II: Deepening Mystery (Sessions 3-4)}

\textbf{Key Scenes}:
\begin{enumerate}[leftmargin=*]
\item Meeting Clan-Mother Yara at her temporary camp
\item Learning about the war-braid and its corruption
\item Investigation of the actual battlefield where the haunting originates
\item Confrontation with ghostly soldiers (combat + social challenges)
\item Discovery that the "music" is the Colonel's tactical commands
\end{enumerate}

\textbf{Clock Advancement}:
\begin{itemize}[leftmargin=*]
\item Singing Wastes Corruption: +2 (haunting spreads to nearby areas)
\item Colonel's Army: +2 (more spirits bound)
\item Clan-Mother's Resolve: +1 (sees ancestors among the enemy)
\end{itemize}

\subsubsection*{Act III: Climax and Resolution (Sessions 5-6)}

\textbf{Key Scenes}:
\begin{enumerate}[leftmargin=*]
\item The final confrontation at the heart of the battlefield
\item Choice: Destroy the war-braid (killing bound spirits) or perform a counter-rite
\item Battle against the Colonel's ghostly army
\item Resolution of the haunting and its consequences
\end{enumerate}

\textbf{Clock Resolution}:
\begin{itemize}[leftmargin=*]
\item If Singing Wastes Corruption fills: The haunting spreads to neighboring regions
\item If Colonel's Army fills: The ghostly force becomes unstoppable without major sacrifice
\item If Clan-Mother's Resolve fills: She performs the ultimate death-rite, becoming part of the haunting
\end{itemize}

\subsection*{7. Resolution Paths}

\begin{description}[leftmargin=*]
\item[Victory (High Cost)] PCs stop the haunting but at great personal/faction cost. The Black Colonel finds peace, but the region is scarred. Award 10-12 XP.
\item[Compromise] Haunting contained but not ended. PCs must return someday. Award 8-10 XP.
\item[Failure] Haunting spreads, more regions affected. PCs flee or are bound themselves. Award 6-8 XP but create ongoing campaign thread.
\item[Noble Failure] PCs lose but save innocents or preserve honor. Award 8-10 XP with strong narrative hook.
\end{description}

\subsection*{8. Custom Mechanics}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{The Singing Mechanic}: When PCs are in the Bone Fields, they must make Wits + Lore rolls (DV 2) to resist the haunting songs. Each failure generates 1 SB that the GM can spend to create haunting effects or advance clocks.

\textbf{War-Braid Bond}: PCs who gain the Clan-Mother's war-braid (Diamond benefit) gain +1 die to social rolls with Ykrul but attract ghostly attention (+1 SB on all rolls while in haunted areas).

\textbf{Echoing Past}: Successful rolls in the ancient battlefield may grant fleeting visions of the original battle, providing useful information but generating 1 SB from temporal dissonance.
\end{mdframed}

\subsection*{9. Campaign Integration}

\begin{itemize}[leftmargin=*]
\item \textbf{Tier I-II}: Introduction to border conflicts and supernatural threats
\item \textbf{Tier III}: Can be revisited when the haunting returns stronger
\item \textbf{Tier IV+}: The Black Colonel may return as a mythic-level threat
\item \textbf{Faction Hooks}: 
\begin{itemize}
\item Ykrul clans now owe the PCs a favor
\item Vilikari noble may offer ongoing employment
\item Ecktorian scholars want reports on the phenomenon
\item Free Companies seek missing comrades
\end{itemize}
\end{itemize}

\section*{Module Creation Quick Reference}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{Step-by-Step Module Creation}:

1. \textbf{Choose a Generator Deck}: Pick the region that fits your theme (Black Banners for war, Mistlands for mystery, Silkstrand for intrigue, etc.)

2. \textbf{Draw Your Seed}: Draw until you have Spade (Place), Heart (Actor), Club (Complication), Diamond (Leverage/Resource)

3. \textbf{Identify the Highest Rank}: This determines your primary clock size and sets the epicenter of your adventure

4. \textbf{Apply Special Rules}: Check for Aces (ongoing omens), Face cards (powerful entities), and regional special mechanics

5. \textbf{Extract Core Elements}:
   \begin{itemize}
   \item What makes this place unique?
   \item Who is the key actor and what do they want?
   \item What's the main complication driving tension?
   \item What resource/leverage can help/hinder the PCs?
   \end{itemize}

6. \textbf{Build the Narrative Hook}: Connect the seed elements into a compelling reason for PCs to care

7. \textbf{Create 2-3 Scene Clocks}: Primary conflict clock, NPC motivation clock, environmental/social pressure clock

8. \textbf{Design Key NPCs}: 
   \begin{itemize}
   \item One antagonist/enigma (from the Heart)
   \item One ally/obstacle (connected to the Club)
   \item One red herring/information source (related to the Spade)
   \end{itemize}

9. \textbf{Plan 3-Act Structure}: Opening discovery, deepening mystery, climax/resolution

10. \textbf{Add Custom Mechanics}: One unique rule that reinforces the theme

11. \textbf{Define Resolution Paths}: Multiple meaningful endings based on PC choices

12. \textbf{Create Campaign Hooks}: How this adventure connects to larger story arcs
\end{mdframed}

\subsection*{Pre-Rolled Adventure Modules}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{Template for Quick Setup}: Each pre-rolled module includes:

\textbf{Header Box}:
\begin{itemize}[leftmargin=*]
\item Adventure Title and Theme
\item Recommended Party Size and Tier
\item Primary Deck Used
\item Estimated Session Length
\end{itemize}

\textbf{Core Elements Section}:
\begin{itemize}[leftmargin=*]
\item The Hook (1-2 paragraphs)
\item Key NPCs with motivations
\item Scene clocks with clear advancement triggers
\item Custom mechanics (if any)
\end{itemize}

\textbf{Scene Cards} (3-5 key encounters):
\begin{itemize}[leftmargin=*]
\item Scene Setup
\item Potential Approaches (combat, social, stealth, etc.)
\item Clock Advancement Opportunities
\item Reward/Penalty Structure
\end{itemize}

\textbf{Resolution Matrix}:
\begin{itemize}[leftmargin=*]
\item Success Conditions
\item Partial Success Outcomes
\item Failure Consequences
\item XP Awards for Different Approaches
\end{itemize}

\textbf{Campaign Integration Notes}:
\begin{itemize}[leftmargin=*]
\item How this fits into larger arcs
\item Seeds for future adventures
\item Faction relationship changes
\end{itemize}
\end{mdframed}

\subsection*{Sample Pre-Rolled Module: "The Merchant's Gamble" (Silkstrand)}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{Header Box}:
\begin{itemize}[leftmargin=*]
\item \textbf{Title}: The Merchant's Gamble
\item \textbf{Theme}: Economic intrigue, family honor, criminal underworld
\item \textbf{Deck}: Silkstrand Generator
\item \textbf{Party}: 3-4 PCs, Tier I-II
\item \textbf{Sessions}: 2-3 sessions
\end{itemize}

\textbf{The Hook}: A prominent Silkstrand merchant family, the Theravels, has hired the PCs to investigate why their spice shipments keep disappearing. What should be a simple escort job becomes complicated when they discover the family's own heir is involved in the thefts.

\textbf{Deck Seed}: 
\begin{itemize}[leftmargin=*]
\item Spade (7 - Spindle Tower with creaking windlass-lifts)
\item Heart (9 - Spinner-matron of the wormhouses)
\item Club (8 - Silk-fungus blight)
\item Diamond (10 - Exchange corner)
\end{itemize}

\textbf{Primary Clock}: \textbf{Exchange Corner Panic Clock} (8 segments) - How long before the market manipulation ruins the Theravel family?

\textbf{Key NPCs}:
\begin{itemize}[leftmargin=*]
\item \textbf{Matron Theravel}: Family matriarch, desperate to save her house's reputation
\item \textbf{Young Theravel}: Reckless heir who's been selling to the Blighters gang
\item \textbf{Spinner-Matron Vessa}: Controls the silk market, orchestrating the manipulation
\end{itemize}

\textbf{Scene Clocks}:
\begin{itemize}[leftmargin=*]
\item Exchange Corner Panic (8): Market destabilization
\item Family Honor (6): Theravel reputation damage  
\item Blighters Gang (6): Criminal organization response
\end{itemize}

\textbf{Resolution Paths}:
\begin{itemize}[leftmargin=*]
\item Expose the heir publicly (family ruined, 6-8 XP)
\item Cover up for the family (market chaos continues, 4-6 XP)
\item Blackmail the Spinner-Matron (wealthy but enemies made, 8-10 XP)
\item Find legitimate solution (honorable but difficult, 10-12 XP)
\end{itemize}
\end{mdframed}

\subsection*{Campaign Adventure Arc: "The Fractured Crown" (Vhasia)}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{Arc Overview}: A 6-8 session campaign using multiple Vhasia draws to create a succession crisis affecting the entire region.

\textbf{Session Structure}:
\begin{enumerate}[leftmargin=*]
\item \textbf{Session 1-2}: "The Empty Throne" - Initial crisis, multiple claimants emerge
\item \textbf{Session 3-4}: "The Noble's Gambit" - PCs choose sides, political intrigue deepens  
\item \textbf{Session 5-6}: "The People's Voice" - Popular uprising, commoners get involved
\item \textbf{Session 7}: "The Military Question" - Army loyalty becomes decisive factor
\item \textbf{Session 8}: "The Coronation" - Final confrontation and resolution
\end{enumerate}

\textbf{Campaign Clocks}:
\begin{itemize}[leftmargin=*]
\item \textbf{Succession Crisis} (10 segments): How close is the kingdom to civil war?
\item \textbf{Noble Conspiracy} (8 segments): How much have the nobles manipulated events?
\item \textbf{Public Unrest} (8 segments): How angry are the common people?
\item \textbf{PC Influence} (6 segments): How much power have the PCs accumulated?
\end{itemize}

\textbf{Major NPCs}:
\begin{itemize}[leftmargin=*]
\item The Lame King's Regents (various factions)
\item Claimant nobles (3-4 main contenders)
\item Popular leaders (merchant guild, labor unions)
\item Military commanders (split loyalties)
\end{itemize}

\textbf{Custom Mechanics}:
\begin{itemize}[leftmargin=*]
\item \textbf{Faction Influence}: PCs can gain influence with different groups through successful rolls
\item \textbf{Public Opinion}: Social rolls affect how the populace views the PCs and claimants
\item \textbf{Military Support}: Certain actions can gain or lose army backing
\end{itemize}

\textbf{Resolution Options}:
\begin{itemize}[leftmargin=*]
\item Support the "legitimate" heir (honorable but politically complex)
\item Back a reformer (risky but potentially transformative)
\item Install a puppet ruler (powerful but creates enemies)
\item Prevent any coronation (chaotic but maintains status quo)
\end{itemize}
\end{mdframed}

\section*{GM Preparation Checklist}

\begin{mdframed}[backgroundcolor=modulebg]
\textbf{Before the Session}:
\begin{itemize}[leftmargin=*]
\item $\square$ Generate adventure seed using appropriate deck
\item $\square$ Create scene clocks with clear advancement triggers
\item $\square$ Prepare key NPCs with motivations and secrets
\item $\square$ Identify potential PC approaches and responses
\item $\square$ Prepare 2-3 Deck of Consequences draws for likely complications
\item $\square$ Set up custom mechanics and explain them clearly
\item $\square$ Plan resolution paths with meaningful differences
\item $\square$ Create campaign integration hooks
\end{itemize}

\textbf{During Play}:
\begin{itemize}[leftmargin=*]
\item $\square$ Announce clocks clearly and update them visibly
\item $\square$ Connect player actions to clock advancement logically
\item $\square$ Offer meaningful choices that affect multiple outcomes
\item $\square$ Let clocks fill when fictionally appropriate
\item $\square$ Provide XP based on engagement and consequences
\end{itemize}
\end{mdframed}

\end{document}
