\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{array}
\usepackage{caption}
\usepackage{changepage}

\pagestyle{fancy}
\fancyhf{}
\rhead{Fate's Edge}
\lhead{Sample Playthrough}
\rfoot{\thepage}

\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

\newcommand{\player}[1]{\textbf{#1:}}
\newcommand{\gm}[1]{\textit{GM: #1}}
\newcommand{\roll}[2]{\texttt{Roll: #1d10 = [#2]}}
\newcommand{\success}[1]{\textbf{Successes: #1}}
\newcommand{\cp}[1]{\textbf{SB: #1}}
\newcommand{\outcome}[1]{\textbf{Outcome: #1}}

\begin{document}

\begin{center}
{\LARGE Sample Playthrough: The Broken Caravan}\\
\vspace{0.5em}
{\large A Fate's Edge Annotated Narrative}
\end{center}

\section*{Session Setup}

\player{GM} Our party consists of Kestra the Arcanist (Wits 4, Arcana 3), Theron the Scout (Wits 3, Survival 3), and Lyra the Blade (Body 4, Melee 3). They're traveling from Ecktoria to Silkstrand through Acasia. Let's set up their first travel leg.

\player{GM} We'll use the Acasia travel deck. I'm drawing until I have all four suits for a Full Seed.

\gm{Draws cards: Spade (6 - Vine-terrace hillside with an abandoned press), Heart (5 - Free Company captain), Club (4 - Bridge feud), Diamond (3 - Monastery letter for bed-and-bread)}

\gm{The highest rank is 6, so we have a 6-segment travel clock. The scene: The party approaches a bridge where two Free Company factions are engaged in a feud. They have a monastery letter that should grant them safe passage.}

\section*{Scene 1: The Bridge Dispute}

\player{GM} As you crest the hill, you see the old stone bridge below, but it's blocked by two groups of mercenaries glaring at each other across the span. One group wears blue tabards, the other red. The bridge keeper looks nervous.

\player{Kestra} I'll try to use my monastery letter to get us safe passage. I'm using Presence + Sway.

\gm{That's a Risky position - they're clearly on edge. Difficulty Value 2.}

\player{Kestra} Rolling Presence 3 + Sway 2 = 5 dice. \roll{5}{8, 6, 4, 1, 1} \success{2} \cp{2}

\outcome{Success \& Cost}

\gm{You approach with the letter, and the blue-tabard captain nods respectfully. "Aye, we honor the brotherhood's seal." But as you start across, the red-tabard leader shouts, "Hold! No one crosses without paying our toll!" He draws his sword. \cp{1} spent: The blue captain steps in front of you protectively, but now both groups are reaching for weapons.}

\player{Theron} I'll scout ahead to see if there's another way across or some advantage we can use. Wits + Stealth.

\gm{Risky position, DV 2. The hillside is rocky but offers some cover.}

\player{Theron} Rolling Wits 3 + Stealth 3 = 6 dice. \roll{6}{9, 7, 5, 3, 1, 1} \success{3} \cp{2}

\outcome{Success \& Cost}

\gm{You slip away and spot a narrow path leading down to the riverbank below - you could ford the stream and climb up on the far side. But \cp{1}: The rocky descent is treacherous. \cp{2}: As you move, you dislodge a small rockslide that alerts both groups to your position.}

\player{Lyra} I say we charge through the middle. I'll lead the way with my blade. Body + Melee.

\gm{Desperate position - you're charging into an armed standoff. DV 3.}

\player{Lyra} Rolling Body 4 + Melee 3 = 7 dice. \roll{7}{10, 8, 6, 4, 3, 1, 1} \success{4} \cp{2}

\outcome{Success \& Cost}

\gm{You leap down onto the bridge, blade singing. The red-tabard captain barely gets his sword up in time. Your fierce charge breaks their morale, and they fall back. But \cp{1}: One of the blue-tabards takes a crossbow bolt meant for you, crying out in pain. \cp{2}: Your charge was so committed you're now surrounded by both groups, with no clear escape route.}

\section*{Scene 2: The Hidden Threat}

\player{GM} As the standoff continues, you notice something odd. The monastery letter in Kestra's hand is... glowing faintly. Strange symbols are appearing on the parchment.

\player{Kestra} I recognize these markings - it's a geomantic ward that activates under stress! I can channel it to help us.

\gm{Intricate action using Wits + Arcana. This is Pressured (DV 2) since you're doing it in combat.}

\player{Kestra} Rolling Wits 4 + Arcana 3 = 7 dice. \roll{7}{9, 7, 6, 4, 3, 1, 1} \success{4} \cp{2}

\outcome{Success \& Cost}

\gm{The letter blazes with light, and the ground beneath both mercenary groups shudders. Cracks spread across the bridge's stones. \cp{1}: The magical backlash leaves you dizzy and off-balance. \cp{2}: The bridge is now unstable - chunks of masonry have fallen into the river.}

\player{Theron} Seeing the bridge weaken, I'll try to convince both groups to retreat before it collapses entirely. Wits + Sway.

\gm{Risky position, DV 2. You're trying to reason with armed men while a bridge crumbles.}

\player{Theron} Rolling Wits 3 + Sway 2 = 5 dice. \roll{5}{7, 6, 4, 2, 1} \success{3} \cp{1}

\outcome{Success \& Cost}

\gm{Your voice carries over the noise: "Do you want to be the last men standing on a broken bridge?" The captains exchange glances. The red-tabard nods sharply and pulls his men back. But \cp{1}: In your focus on the mercenaries, you didn't notice the bridge's central support was already compromised.}

\section*{Scene 3: The Escape}

\player{GM} With a grinding roar, the bridge's center span begins to collapse. You have seconds to get clear.

\player{Lyra} I'll help the wounded blue-tabard back from the edge. Body + Athletics.

\gm{Pressured action, DV 2. You're carrying extra weight while the ground shakes.}

\player{Lyra} Rolling Body 4 + Athletics 2 = 6 dice. \roll{6}{8, 6, 5, 3, 1, 1} \success{3} \cp{2}

\outcome{Success \& Cost}

\gm{You grab the wounded man and haul him toward the shore just as the bridge span crashes into the river below. He's alive but injured. \cp{1}: The effort strains your shoulder. \cp{2}: As you reach safety, you realize the collapse has blocked the main road completely.}

\player{Kestra} I'll use my geomancy to assess if there's a safe path through the rubble. Wits + Survival.

\gm{Controlled position (you're safely ashore), DV 1.}

\player{Kestra} Rolling Wits 4 + Survival 1 = 5 dice. \roll{5}{9, 6, 4, 3, 2} \success{3} \cp{0}

\outcome{Clean Success}

\gm{You spot a gap in the rubble where the old press's foundation created a natural tunnel. It's a tight squeeze, but it will let you continue toward Silkstrand.}

\section*{Travel Resolution}

\player{GM} Let's check our travel clock. We had 6 segments. We've resolved the major conflict, found an alternate route, and survived the encounter.

\gm{I'll tick 2 segments for successfully navigating the obstacle. The party earned 1 Boon for creative problem-solving and 2 XP for the Complication Dividend (face cards drawn).}

\section*{GM Reflection}

\gm{This scene demonstrates several key Fate's Edge principles:}

\begin{itemize}[leftmargin=*]
\item \textbf{Risk as Drama}: Every successful action still generated complications through SB.
\item \textbf{Player Agency}: The players had multiple approaches (diplomacy, stealth, direct action) that all contributed meaningfully.
\item \textbf{Narrative Primacy}: The collapsing bridge wasn't just a mechanical effect - it changed the story and forced new decisions.
\item \textbf{Meaningful Growth}: XP was earned through engagement with complications, not just success.
\item \textbf{Collaborative Storytelling}: The glowing letter was a player detail that I incorporated into the fiction to create new possibilities.
\end{itemize}

\gm{Key mechanics shown:}
\begin{itemize}[leftmargin=*]
\item Outcome Matrix application (Success \& Cost was the most common result)
\item Story Beat spending to add tension and story elements
\item Description Ladder use (Intricate action for Kestra's geomancy)
\item Travel clock advancement based on resolution quality
\end{itemize}

\gm{The session ended with the party making progress but also creating new story threads: an injured NPC to potentially care for, a blocked road that might affect future travelers, and questions about the magical monastery letter.}

\end{document}
