\documentclass[12pt]{article}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{tikz}
\usepackage{framed}

\titleformat{\section}{\large\bfseries}{\thesection.}{1em}{}
\titleformat{\subsection}{\bfseries}{\thesubsection.}{1em}{}
\titleformat{\subsubsection}{\bfseries\itshape}{\thesubsubsection.}{1em}{}

\title{Assets for Fate's Edge}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Assets}

Assets in Fate's Edge are off-screen resources that extend your influence beyond personal capability. They solve problems between sessions and provide leverage when you return to the table. Assets do not act in scenes directly, but they change the fiction and provide leverage. All Assets have a Condition Track: Maintained, Neglected, or Compromised.

\subsection*{Minor Assets (4 XP, 1 day)}

\begin{description}[leftmargin=*]
  \item[\textbf{Signature Weapon}] Named heirloom or masterpiece. Off-Screen: Mitigate gear-break CP 1; recovery scene clears Compromised. \\
  \textbf{Scene Keys:} Personal combat, duels, ceremonial displays \\
  \textbf{Hooks:} Weapon demands tribute; Rival claims it's theirs; Ancient curse stirs.

  \item[\textbf{Scholar's Cell}] Quiet room with texts and desk. Off-Screen: One research opener starts Controlled or +1 die. \\
  \textbf{Scene Keys:} Libraries, monasteries, private study \\
  \textbf{Hooks:} Forbidden tome sought by others; Research attracts unwanted attention.

  \item[\textbf{Safehouse Network}] Hidden bolt-holes in one city. Off-Screen: Clear Hunt 1 between scenes or start Controlled for entry. \\
  \textbf{Scene Keys:} Urban environments, criminal districts \\
  \textbf{Hooks:} Safehouse compromised; Owner demands favor; New location needed.

  \item[\textbf{Artisan's Workshop}] Tools + bench for field repairs. Off-Screen: Clear Compromised from one item. \\
  \textbf{Scene Keys:} Craft districts, frontier outposts \\
  \textbf{Hooks:} Rare materials needed; Rival artisan challenges you; Fire threatens workshop.

  \item[\textbf{Courier Network}] Trusty runners and dead-drops. Off-Screen: Deliver message; next parley starts Controlled or Mitigate CP 1. \\
  \textbf{Scene Keys:} Trade routes, noble courts \\
  \textbf{Hooks:} Courier captured; Message intercepted; Network disrupted by war.

  \item[\textbf{Small Farm/Orchard}] Larder and surplus. Off-Screen: Waive minor Supply complication or refresh "consumable" gear tag. \\
  \textbf{Scene Keys:} Rural areas, family holdings \\
  \textbf{Hooks:} Blight threatens crops; Raiders attack; Harvest festival draws attention.

  \item[\textbf{Local Tavern/Inn}] Neighborhood hub. Off-Screen: First social opener starts Controlled. \\
  \textbf{Scene Keys:} Market towns, crossroads \\
  \textbf{Hooks:} New owner changes rules; Regular patrons in trouble; Noble guests cause issues.

  \item[\textbf{Minor Shrine/Temple}] Hearth for the spirit. Off-Screen: Clear Fatigue 1 or Mitigate first morale CP 1. \\
  \textbf{Scene Keys:} Religious sites, places of pilgrimage \\
  \textbf{Hooks:} Heresy investigated; Sacred relic stolen; Pilgrims bring trouble.

  \item[\textbf{Fishing Boat/Barge}] River/harbor skiff. Off-Screen: Water crossing starts Controlled or ignore first travel Supply loss. \\
  \textbf{Scene Keys:} Waterways, coastal areas \\
  \textbf{Hooks:} Boat requires repairs; Storm damages vessel; Smugglers use similar craft.

  \item[\textbf{Herbal Garden}] Beds of restorative greens. Off-Screen: Remove Fatigue 1 from one character. \\
  \textbf{Scene Keys:} Rural estates, healer's abodes \\
  \textbf{Hooks:} Rare herb needed; Garden raided; Poisonous plant misidentified.
\end{description}

\subsection*{Standard Assets (8 XP, 1 week)}

\begin{description}[leftmargin=*]
  \item[\textbf{Noble Title/Charter}] Recognized rank and privileges. Off-Screen: Secure audience; first petition starts Controlled or Protected Turn. \\
  \textbf{Scene Keys:} Noble courts, administrative centers \\
  \textbf{Hooks:} Title challenged; Political rival schemes; Duties demand attention.

  \item[\textbf{Guild Membership}] Standing in a working order. Off-Screen: Procurement/commission starts Controlled or Mitigate CP 1. \\
  \textbf{Scene Keys:} Guild halls, trade districts \\
  \textbf{Hooks:} Guild politics complicate matters; Rival guild interferes; Master demands tribute.

  \item[\textbf{Merchant Stall/Front}] Fixed face in the market. Off-Screen: Free upkeep on one Minor Asset or +1 die on barter. \\
  \textbf{Scene Keys:} Marketplaces, trade fairs \\
  \textbf{Hooks:} Stall fire destroys inventory; Competitor undercuts prices; Tax collector visits.

  \item[\textbf{Spy Ring/Informants}] Eyes and ears on retainer. Off-Screen: Targeted inquiry starts Controlled or +1 die; GM adds social Hook. \\
  \textbf{Scene Keys:} Urban intrigue, political centers \\
  \textbf{Hooks:} Informant compromised; False information planted; Rival network discovered.

  \item[\textbf{Mercenary Contract}] Reach for hired steel. Off-Screen: Introduce temporary Cap 2 unit for one beat (consumes Initiative). \\
  \textbf{Scene Keys:} Battlefields, frontier conflicts \\
  \textbf{Hooks:} Contract terms disputed; Mercenaries demand more pay; Unit suffers losses.

  \item[\textbf{Trading Charter}] Right of trade along routes. Off-Screen: At customs/gates, Mitigate first Diamonds/Clubs CP 1 or start Controlled. \\
  \textbf{Scene Keys:} Trade routes, border crossings \\
  \textbf{Hooks:} Charter revoked; New taxes imposed; Rival merchant challenges rights.

  \item[\textbf{Library Archive}] Stacks and custodian. Off-Screen: Deep topic research starts Controlled and +1 die. \\
  \textbf{Scene Keys:} Centers of learning, ancient sites \\
  \textbf{Hooks:} Rare manuscript sought by others; Archive damaged by fire; Forbidden knowledge uncovered.

  \item[\textbf{Craftsman's Quarter}] Shop + apprentices. Off-Screen: Produce quality piece: clear Compromised or grant +1 die once. \\
  \textbf{Scene Keys:} Craft districts, industrial areas \\
  \textbf{Hooks:} Master craftsman dies; Apprentices cause trouble; Rare materials needed.

  \item[\textbf{Temple Complex}] Choirs, courts, care. Off-Screen: Among faithful, gain Protected Turn or Mitigate first Sanctity/Crowd CP 1. \\
  \textbf{Scene Keys:} Religious centers, pilgrimage sites \\
  \textbf{Hooks:} Heresy investigated; Sacred relic stolen; Pilgrims bring trouble.

  \item[\textbf{Port Authority}] Berths, booms, bells. Off-Screen: Priority handling: Mitigate first Curfew/Crowd CP 1 or loading starts Controlled. \\
  \textbf{Scene Keys:} Harbor cities, maritime centers \\
  \textbf{Hooks:} Quarantine declared; Customs officials corrupt; Naval blockade threatens.
\end{description}

\subsection*{Major Assets (12 XP, 1 month)}

\begin{description}[leftmargin=*]
  \item[\textbf{City District License}] You run a ward's levers. Off-Screen: In-district, civic opener starts Controlled or set local rail 1. \\
  \textbf{Scene Keys:} Urban districts, administrative centers \\
  \textbf{Hooks:} Riots threaten district; Political rival schemes; Imperial audit demanded.

  \item[\textbf{Regional Monopoly}] Exclusive right in a staple. Off-Screen: In market scenes, cancel first Diamonds CP. \\
  \textbf{Scene Keys:} Trade centers, resource regions \\
  \textbf{Hooks:} Competitors emerge; Supply disrupted; Monopoly investigated.

  \item[\textbf{Fortress/Stronghold}] A hard place held by your banner. Off-Screen: In siege/raid, ignore first Spades complication. \\
  \textbf{Scene Keys:} Strategic locations, frontier posts \\
  \textbf{Hooks:} Siege engines deployed; Treason within walls; Supplies run low.

  \item[\textbf{Large Estate/Demesne}] Fields, mills, partys. Off-Screen: No minor Supply complications or free upkeep on one additional Asset. \\
  \textbf{Scene Keys:} Rural estates, agricultural regions \\
  \textbf{Hooks:} Peasant revolt threatens; Harvest failure imminent; Rival noble presses claims.

  \item[\textbf{Naval Fleet Share}] Oars, sails, signal flags. Off-Screen: Ensure safe passage: maritime travel starts Controlled and Mitigate CP 1. \\
  \textbf{Scene Keys:} Coastal areas, naval bases \\
  \textbf{Hooks:} Mutiny aboard ships; Naval battle imminent; Storm damages fleet.

  \item[\textbf{University College}] Faculty, quad, seal. Off-Screen: Downtime reduces one Skill advance cost by 1 (min 1). \\
  \textbf{Scene Keys:} Centers of learning, academic districts \\
  \textbf{Hooks:} Student protests disrupt classes; Rival institution challenges; Endowment threatened.

  \item[\textbf{Banking Charter}] Letters, ledgers, liens. Off-Screen: In finance scenes, Mitigate first Diamonds CP 1; if clean, +1 die. \\
  \textbf{Scene Keys:} Financial districts, merchant centers \\
  \textbf{Hooks:} Audit reveals discrepancies; Rivals attempt bankruptcy; Currency devalued.

  \item[\textbf{Diplomatic Enclave}] Flag, immunity, ink. Off-Screen: In foreign courts, ignore first Clubs complication or parley starts Controlled. \\
  \textbf{Scene Keys:} Foreign capitals, diplomatic quarters \\
  \textbf{Hooks:} Diplomatic incident occurs; Immunity challenged; War threatens relations.

  \item[\textbf{Mine/Quarry Operation}] Raw stone, hard people. Off-Screen: Provide materials: clear Compromised on structure/gear or +1 die once. \\
  \textbf{Scene Keys:} Mining regions, industrial areas \\
  \textbf{Hooks:} Cave-in threatens workers; Rare mineral discovered; Labor strike organized.

  \item[\textbf{Market Control}] You set the bell. Off-Screen: Declare Surplus/Scarcity for staple; negotiation about it starts Controlled. \\
  \textbf{Scene Keys:} Major marketplaces, trade centers \\
  \textbf{Hooks:} Competitors undercut prices; Supply disrupted; Public riot over prices.
\end{description}

\subsection*{Tier III Assets (16-20 XP, 3 months)}

\begin{description}[leftmargin=*]
  \item[\textbf{Grand Forge--Citadel}] City-sized heat and hammer. Off-Screen: Produce masterworks; ignore first gear complication or clear batch Compromised. \\
  \textbf{Scene Keys:} Dwarven holds, industrial centers \\
  \textbf{Hooks:} Forge-fires threaten to burn out; Ancient techniques rediscovered; Rival holds compete.

  \item[\textbf{Harbor Light Consortium}] Beacons from cape to cape. Off-Screen: Dictate right-of-way: ally's maneuver starts Controlled or 10-ft zone Hazard 1. \\
  \textbf{Scene Keys:} Maritime routes, coastal regions \\
  \textbf{Hooks:} Beacon chain compromised; Naval battle disrupts signals; Storm damages lights.

  \item[\textbf{Grand Exchange Charter}] Markets move when you breathe. Off-Screen: In halls, cancel one Diamonds CP; if clean, +1 die. \\
  \textbf{Scene Keys:} Major trade centers, financial districts \\
  \textbf{Hooks:} Market crash threatens economy; Rival exchange opens; Trade war erupts.

  \item[\textbf{Thepyrgosi Great Library}] The north's memory palace. Off-Screen: Major inquiry starts Controlled and +1 die; Mitigate first scholarly CP 1. \\
  \textbf{Scene Keys:} Centers of learning, ancient sites \\
  \textbf{Hooks:} Rare manuscript sought by others; Archive damaged by fire; Forbidden knowledge uncovered.

  \item[\textbf{Kahfagian Fleet Command}] Signals across sea lanes. Off-Screen: Declare blockade/corridor: hostile exits Escape +1 or friendly passage Controlled. \\
  \textbf{Scene Keys:} Naval bases, maritime regions \\
  \textbf{Hooks:} Naval mutiny threatens control; Enemy fleet approaches; Storm scatters ships.

  \item[\textbf{Oshiiran Hydraulic Empire}] Water is wealth. Off-Screen: Set Irrigated tag for route; cancel first Supply/shortage Diamonds CP. \\
  \textbf{Scene Keys:} River valleys, agricultural regions \\
  \textbf{Hooks:} Dam threatened by earthquake; Canal sabotage disrupts flow; Drought affects entire region.

  \item[\textbf{Ashaani Palace Complex}] Keys to office and omen. Off-Screen: Political opener starts Controlled and gains Protected Turn. \\
  \textbf{Scene Keys:} Noble courts, administrative centers \\
  \textbf{Hooks:} Palace coup attempted; Ancient curse disturbs nobles; Foreign embassy causes incident.

  \item[\textbf{Valewood Grove Sanctuary}] Old roots, older vows. Off-Screen: Wilderness scenes, cancel natural Hazard tick or Mitigate CP 1. \\
  \textbf{Scene Keys:} Ancient forests, mystical sites \\
  \textbf{Hooks:} Forest spirits angered; Loggers threaten grove; Ancient entity awakens.

  \item[\textbf{Dwarven Mountain Hall}] Deep law, deeper craft. Off-Screen: Engineering/crafting opener starts Controlled and +1 die. \\
  \textbf{Scene Keys:} Mountain holds, underground cities \\
  \textbf{Hooks:} Cave-in threatens tunnels; Ancient forge reactivated; Rival clan disputes territory.

  \item[\textbf{Ykrul Steppe Confederation}] Ten thousand hooves. Off-Screen: On steppe, group movement starts Controlled and Escape 1. \\
  \textbf{Scene Keys:} Grasslands, nomad territories \\
  \textbf{Hooks:} Clan war threatens unity; Harsh winter approaches; Foreign army invades steppe.
\end{description}

\subsection*{Specialized Assets (Variable XP)}

\begin{description}[leftmargin=*]
  \item[\textbf{Magical Laboratory} (10 XP)] Arcane benches, wards. Off-Screen: Ritual Backlash 1 CP once; experiment starts Controlled. \\
  \textbf{Scene Keys:} Wizard towers, arcane institutions \\
  \textbf{Hooks:} Experiment goes wrong; Rival mage steals research; Magical contamination spreads.

  \item[\textbf{Healing Sanctuary} (8 XP)] Beds, herbs, bells. Off-Screen: Treat patient: clear Harm 1 or Fatigue 2 (downtime only). \\
  \textbf{Scene Keys:} Religious sites, medical centers \\
  \textbf{Hooks:} Plague overwhelms sanctuary; Sacred relic stolen; Heretical healing method discovered.

  \item[\textbf{Performance Venue} (6 XP)] Stage + crowd. Off-Screen: Build reputation; first Performance opener starts Controlled or +1 die. \\
  \textbf{Scene Keys:} Theaters, entertainment districts \\
  \textbf{Hooks:} Rival performer challenges you; Audience riot threatens venue; Noble patron demands attention.

  \item[\textbf{Observatory/Star Tower} (8 XP)] Ephemerides and lenses. Off-Screen: Predict timing; navigation/timing opener starts Controlled. \\
  \textbf{Scene Keys:} High peaks, centers of learning \\
  \textbf{Hooks:} Celestial event disrupts predictions; Rival observatory claims discovery; Stars themselves change.

  \item[\textbf{Alchemical Garden} (10 XP)] Beds of rare reagents. Off-Screen: Produce draught; once/session Mitigate first toxin CP 1 or +1 die salve. \\
  \textbf{Scene Keys:} Remote estates, mystical locations \\
  \textbf{Hooks:} Rare plant needed for crucial potion; Garden raided by thieves; Poisonous cross-pollination occurs.

  \item[\textbf{Training Academy} (12 XP)] Masters, drills, mats. Off-Screen: Downtime reduces one Skill advance cost by 1 (min 1) for one trainee. \\
  \textbf{Scene Keys:} Military bases, educational institutions \\
  \textbf{Hooks:} Instructor scandal threatens reputation; Student rebellion disrupts classes; Rival academy challenges methods.

  \item[\textbf{Intelligence Bureau} (10 XP)] Analysts and cut-outs. Off-Screen: Avoid surprise; first recon opener starts Controlled. \\
  \textbf{Scene Keys:} Government buildings, urban centers \\
  \textbf{Hooks:} Agent compromised; False intelligence planted; Rival network discovered.

  \item[\textbf{Hospitality House} (6 XP)] Beds, bread, rumor. Off-Screen: Gain Protected Turn in hosted parley or Mitigate first Crowd CP 1. \\
  \textbf{Scene Keys:} Crossroads, neutral territories \\
  \textbf{Hooks:} Unwelcome guest arrives; Food shortage threatens hospitality; Rival house challenges your rights.

  \item[\textbf{Storage Vault} (6 XP)] Locks, ledgers, layers. Off-Screen: Ignore first theft/sabotage complication targeting stored goods. \\
  \textbf{Scene Keys:} Commercial districts, secure facilities \\
  \textbf{Hooks:} Vault breached by thieves; Rare item demanded by authorities; Fire threatens stored goods.

  \item[\textbf{Communication Relay} (8 XP)] Fires, flags, drums. Off-Screen: Coordinate afar: timed coordination starts Controlled or Mitigate CP 1. \\
  \textbf{Scene Keys:} Strategic locations, military posts \\
  \textbf{Hooks:} Relay station attacked; Message intercepted; System overwhelmed by traffic.
\end{description}

\subsection*{Asset Upgrades and Variants (+2-4 XP)}

Enhance your Assets with additional investments:

\begin{description}[leftmargin=*]
  \item[\textbf{Enhanced Security} (+2 XP)] Hardened doors, better watchers. Effect: Mitigate first theft/sabotage CP 1 each session.

  \item[\textbf{Expanded Scope} (+3 XP)] Wider footprint. Effect: Use Off-Screen effect twice/session or across two adjacent venues.

  \item[\textbf{Magical Enhancement} (+4 XP)] Warded, sanctified. Effect: Add Mitigate Backlash 1 (once/session) for rituals OR Hazard 1 aura (10-ft, 1 beat).

  \item[\textbf{Political Connections} (+3 XP)] Papers, patrons. Effect: In civic/authority scenes, Mitigate first Clubs CP 1.
\end{description}

\subsection*{Asset Condition States}

All Assets track their condition through narrative and mechanical states:

\begin{description}[leftmargin=*]
  \item[\textbf{Maintained}] Full capability; no penalties. Asset functions as described.

  \item[\textbf{Neglected}] -1 die when used; requires attention. Narrative: slower, sullen, short-staffed. \\
  \textbf{Recovery:} Spend downtime repairing/servicing it, or pay XP immediately for replacement parts.

  \item[\textbf{Compromised}] Unavailable. Narrative: captured, burned, seized, defected. \\
  \textbf{Recovery:} Requires significant downtime, resources, or a narrative quest to restore.
\end{description}

\subsection*{Using Assets}

Assets provide off-screen benefits:
\begin{itemize}
  \item Use each Asset's listed Off-Screen effect once per session for free.
  \item To reshape the current scene, spend 1 Boon.
  \item The Asset must have scope and reach.
  \item Superior and Artifact gear do not require upkeep, but if Compromised through Complications, only narrative quests can repair them.
\end{itemize}

\subsection*{Asset Creation Guidelines}

To create a custom Asset:
\begin{enumerate}
  \item \textbf{Determine Tier} — Minor (4 XP), Standard (8 XP), Major (12 XP), or Tier III (16-20 XP)
  \item \textbf{Define Off-Screen Effect} — What clear, codified outcome does it provide?
  \item \textbf{Set Scene Keys} — Where does it naturally provide advantage?
  \item \textbf{Create Hooks} — What complications can arise from owning it?
  \item \textbf{Consider Upgrades} — What enhancements make sense?
  \item \textbf{Balance Scope} — Ensure the effect is plausible and appropriately powerful for its tier.
\end{enumerate}

\end{document}
