\documentclass[12pt]{article}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning}
\usepackage{framed}

\titleformat{\section}{łarge\bfseries}{þesection.}{1em}{}
\titleformat{\subsection}{\bfseries}{þesubsection.}{1em}{}
\titleformat{⫕section}{\bfseriesℹtshape}{þesubsubsection.}{1em}{}

\title{Backgrounds and Prestige Backgrounds for Fate's Edge}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Backgrounds and Prestige Backgrounds}

Backgrounds in Fate's Edge are narrative packages that tie your character to a specific place, faction, or tradition. Each requires ownership of a specific Asset and grants a Talent, Perk, and narrative hooks that shape how you engage with the world. Prestige Backgrounds are epic−tier roles earned in play, with powerful abilities and world−shaping consequences.

\subsection*{Regional Backgrounds}

\begin{description}[leftmargin=*]
  ℹtem[\textbf{Vhasian Gentry} (Asset Required: Vhasian Manor, Estate, or County)]
    \textbf{Talent (4 XP):} Influence at Court (Vhasian) −−− +1 die when making formal petitions and protocol interactions inside Vhasian courts and councils. \\
    \textbf{Perk (Choose 1/session):} +1 die on one opener in a keyed venue; Hazard 1 for one beat in a named environment; Mitigate first SB for on−theme twists; Restore 1 follower (clear 1 Exposure or heal Harm 1); Free upkeep on one owned Asset. \\
    \textbf{Scene Keys:} Court/Audience, Noble Reception \\
    \textbf{Hooks (GM on SB spend):} A rival house "assists" by rerouting your petition; Protocol demands a costly courtesy now; A jealous clerk bumps another case ahead. \\
    \textbf{Invite (optional):} "Defer to protocol before acting."

  ℹtem[\textbf{Aeler Tower Savant} (Asset Required: Aeler Vault−Citadel or Scholar's Tower)]
    \textbf{Talent (5 XP):} Deep Study (Aeler) −−− Once per session, when researching ancient or forbidden lore, gain start Controlled or +1 die. \\
    \textbf{Perk (Choose 1/session):} Gain one free Boon for arcane research; Mitigate first SB from magical Backlash this scene; +1 die to one Arcana roll; Access restricted vault section once. \\
    \textbf{Scene Keys:} Deep Archive Stacks, Forbidden Laboratories \\
    \textbf{Hooks (GM on SB spend):} Ancient guardian awakens; Forbidden knowledge carries a curse; Rival scholar sabotages your work. \\
    \textbf{Invite (optional):} "Quote the Deep Codices before acting."

  ℹtem[\textbf{Mistlands Warden} (Asset Required: Bell−Line Levee or Protectorate Commission)]
    \textbf{Talent (6 XP):} Breath−Test (Mistlands) −−− Once per session, perform the Aeler rite to diagnose a breach: name the incursion type and approximate strength. \\
    \textbf{Perk (Choose 1/session):} Command a Cap 3 Protectorate squad; Sanctify a small area with ward−salt and bells; Decode bell−logs to know quiet/hot watches. \\
    \textbf{Scene Keys:} Weeping Gate, Bell−Line Levee \\
    \textbf{Hooks (GM on SB spend):} Local Mistlander challenges Aeler authority; New horror exceeds protocol; Procedure demands sacrifice. \\
    \textbf{Invite (optional):} "Trust the bells before acting."

  ℹtem[\textbf{Silkstrand Span−Factor} (Asset Required: Bridge Charter or Span−Office)]
    \textbf{Talent (5 XP):} Bridge Authority (Silkstrand) −−− +1 die when negotiating passage rights or dealing with bridge guilds. \\
    \textbf{Perk (Choose 1/session):} Priority crossing on any bridge; +1 die to Hazard rolls involving bridges; Waive one bridge toll this scene; Gain start Controlled for bridge inspections. \\
    \textbf{Scene Keys:} Three−Queens Bridge, Redwater Dyeworks \\
    \textbf{Hooks (GM on SB spend):} Bridge riot blocks passage; Counterfeit bridge rights discovered; Structural weakness requires immediate attention. \\
    \textbf{Invite (optional):} "Check the span registers before crossing."

  ℹtem[\textbf{Kahfagian Spur Captain} (Asset Required: Spur Station or Relay Charter)]
    \textbf{Talent (5 XP):} Lantern−Law Authority (Kahfagia) −−− +1 die when invoking lantern−law or dealing with maritime authorities. \\
    \textbf{Perk (Choose 1/session):} Predict next beacon shift; Gain start Controlled for pilotage actions; Halve customs levy through flawless paperwork; Minor port official backs your claim. \\
    \textbf{Scene Keys:} Mirror−Light Tower, Admiralty Quay \\
    \textbf{Hooks (GM on SB spend):} Lanterns malfunction during critical passage; Rival cites deeper bylaw; Beacon chain breaks during storm. \\
    \textbf{Invite (optional):} "Cite chapter and verse of lantern−law."

\end{description}

\subsection*{Cultural Backgrounds}

\begin{description}[leftmargin=*]
  ℹtem[\textbf{Ykrul Moor Warden} (Asset Required: Clan Steading or Boundary Mark)]
    \textbf{Talent (6 XP):} Blood−Sense (Ykrul) −−− Once per session, sense the emotional state of a named person within one day's ride. \\
    \textbf{Perk (Choose 1/session):} +1 die to tracking a specific individual; Call upon clan hospitality once; Gain start Controlled in steppe negotiations; Mitigate first SB from weather this scene. \\
    \textbf{Scene Keys:} Wolf Road Milepost, Clan Gathering \\
    \textbf{Hooks (GM on SB spend):} Clan feud rekindles; Rasputitsa traps your party; Salt shortage creates tension. \\
    \textbf{Invite (optional):} "Swear by the wolf before acting."

  ℹtem[\textbf{Aelaerem Hearth−Keeper} (Asset Required: Family Homestead or Moot Oak Rights)]
    \textbf{Talent (5 XP):} Hearth Sense (Aelaerem) −−− Once per session, know if a named person is in genuine need or danger within your region. \\
    \textbf{Perk (Choose 1/session):} Guest−loaf and salt for safe board; +1 die to Mediation rolls; Gain start Controlled for local social actions; Red thread ties a minor feud. \\
    \textbf{Scene Keys:} Moot Oak, Cider−Press Barn \\
    \textbf{Hooks (GM on SB spend):} Scarecrow turns hostile; Soured wassail creates ill−will; Chalk maze fills with mist. \\
    \textbf{Invite (optional):} "Tie the red thread before speaking."

  ℹtem[\textbf{Tulkani Pit−Runner} (Asset Required: Pit Charter or Mask Collection)]
    \textbf{Talent (6 XP):} Shadow−Cant (Tulkani) −−− Once per session, communicate secretly with any Tulkani within earshot. \\
    \textbf{Perk (Choose 1/session):} +1 die to Performance in disguise; Gain start Controlled for covert operations; Mitigate first SB from social detection; Free upkeep on one Shadow Asset. \\
    \textbf{Scene Keys:} Crow's Roost, Masked Gatherings \\
    \textbf{Hooks (GM on SB spend):} Performance judged heretical; Shadow−cant intercepted; Guild challenges your pit rights. \\
    \textbf{Invite (optional):} "Don the mask before speaking."

\end{description}

\subsection*{Prestige Backgrounds (Epic Tier)}

\begin{description}[leftmargin=*]
  
  ℹtem[\textbf{The Cursed Child of Silkstrand}] 
  \textbf{Requires:} Personally mitigate or trigger a major Acasian Curse event; deep tie to Silkstrand. \\
  \textbf{Talent (20 XP, 1/campaign):} Laughter That Ends Sieges −−− In a moment of absolute desperation, focus the Curse through yourself. End one large−scale conflict in sight (siege, battle, storm) as reality unravels it. No roll. \\
  \textbf{Epic Hook −−− The Curse Hungers:} The Curse binds to you. At the start of every session, the GM introduces a minor, personal echo (an omission, loop, or wrong reflection) affecting you or allies. GM: bank +2 SB and start/advance a 6−segment Silkstrand Echo Front. \\
  \textbf{Deck Hooks:} Treat any Valewood/Theona 9 drawn in your scenes as yours. Spending a local (e.g., Way−cord) can postpone, not erase, an echo. \\
  \textbf{Scene Keys:} Cursed Crossroads, Affected Estate \\
  \textbf{Hooks (GM on SB spend):} The Curse twists a future success into a failure; A cult of the Curse marks you for meddling; Break one knot; two tighten elsewhere.

  ℹtem[\textbf{Stone−Speaker of Khaz−Vurim}] 
  \textbf{Requires:} Broker a monumental peace between Aeler holds or with a major power; dwarven culture or unprecedented trust. \\
  \textbf{Talent (20 XP, 1/campaign):} Move the Mountain's Will −−− Call the deep will of the Aelerians to reroute a geological or political reality: open a pass weeks early, flood a rival shaft, or cause a hold to ignore a king's edict. No roll. \\
  \textbf{Epic Hook −−− The Mountain's Price:} You swear a Geas of Stone. The GM assigns a monumental, slow task against your interests. Until completed, you lose dwarven hospitality and stone−magic benefit (treat social Position as Desperate with Aeler guilds). \\
  \textbf{Deck Hooks:} When used, flip any Aeler played this arc immediately. Start/advance an 8−segment Geas of Stone Front. \\
  \textbf{Scene Keys:} Royal Crypt, Under−Vault Archive \\
  \textbf{Hooks (GM on SB spend):} Ancestral spirit misidentifies you; Senior Adept deems your rite overreaching; You learn a truth damning to a living house.

  ℹtem[\textbf{The Last Dauphin's Shadow}] 
  \textbf{Requires:} Publicly champion, protect, or expose the Sun Court claimant; irrevocably entangled in the cause. \\
  \textbf{Talent (18 XP, 1/campaign):} The Sunburst Warrant −−− Issue commands obeyed as if from the lost Sun−King: mobilize hidden cells, muster a county levy, or claim noble sanctuary. Choose one concrete objective; it happens, unless physically impossible. No roll. \\
  \textbf{Epic Hook −−− The Pretender's Toll:} Each use consumes hope. GM selects a loyalist asset to be lost, disillusioned, or exposed. Start/advance a 6−segment Candle of the Sun Front; when it fills, the movement fractures. \\
  \textbf{Deck Hooks:} In Vhasia, treat face−card s as predisposed but legal/economic hits escalate faster. \\
  \textbf{Scene Keys:} Sun Palace, Parlement Hall \\
  \textbf{Hooks (GM on SB spend):} Rival cites deeper bylaw; Your precedent is weaponized immediately; Lanterns malfunction, proving you right but creating Hazard.

  ℹtem[\textbf{Mist−Seer of the Weeping Gate}] 
  \textbf{Requires:} Endure the Direwood and return changed; face a horror beyond common wraiths. \\
  \textbf{Talent (20 XP, 1/campaign):} Bargain with the Deep Mist −−− Open parley with the Mist's sentience. Ask one true question or command it to withhold influence from a named place/person for a year and a day. No roll. \\
  \textbf{Epic Hook −−− The Wood Remembers:} Pay a Bargain of Breath: surrender a fundamental memory/name/sense/skill (permanent). GM banks +2 SB and starts a 6−segment Lost Thing Front keyed to what was given up. \\
  \textbf{Deck Hooks:} A Mistlands (Ward−salt or Lantern Writ) can soften−−−not negate−−−the next echo of the loss. \\
  \textbf{Scene Keys:} Direwood Edge, Pall Watch−tower \\
  \textbf{Hooks (GM on SB spend):} Local Mistlander challenges Aeler authority; New horror exceeds protocol; Procedure demands sacrifice.

  ℹtem[\textbf{Keeper of the Foedus Stone}] 
  \textbf{Requires:} Primary mediator preventing a Ykrul/Vilikari (or equivalent) war from breaking. \\
  \textbf{Talent (18 XP, 1/campaign):} The Law That Binds Spears −−− At the Foedus Stone (or an ancient treaty site), impose an absolute truce between two named factions. For its span, any violent act across the line triggers immediate supernatural calamity on the offender. No roll. \\
  \textbf{Epic Hook −−− The Balance of Grudges:} The hate is compressed, not erased. Begin an 8−segment Coiled Blood Front that passively ticks each session. When it blows, the rematch is worse−−−your name is on the blame. \\
  \textbf{Deck Hooks:} During the truce, s like Foedus Seal gain +1 practical scope but any hostage automatically targets your allies first on SB spend. \\
  \textbf{Scene Keys:} Foedus Stone, Clan Gathering \\
  \textbf{Hooks (GM on SB spend):} Clan feud rekindles; Rasputitsa traps your party; Salt shortage creates tension.

\end{description}

\subsection*{Background Creation Guidelines}

To create a custom Background:
\begin{enumerate}
  ℹtem \textbf{Name It} −−− Give it an evocative title.
  ℹtem \textbf{Asset Requirement} −−− Tie it to a Minor/Standard/Major Asset.
  ℹtem \textbf{Talent} −−− Choose or design a thematically appropriate Talent (3−6 XP).
  ℹtem \textbf{Perk} −−− Pick one mechanical benefit (session−limited) with 2−3 options.
  ℹtem \textbf{Scene Keys} −−− Name two venues where it naturally appears.
  ℹtem \textbf{Hooks} −−− List 2−−3 SB sinks with narrative prompts.
  ℹtem \textbf{Invite (optional)} −−− Behavior cue that may earn a Boon.
\end{enumerate}

Prestige Backgrounds follow the same structure but require:
\begin{itemize}
  ℹtem Epic−tier prerequisites (significant story investment)
  ℹtem Once−per−campaign Talents (12−20 XP)
  ℹtem Epic Hooks that create lasting world changes
  ℹtem Integration with campaign timers and fronts
\end{itemize}

\end{document}
