\documentclass[12pt]{article}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}

\titleformat{\section}{\large\bfseries}{\thesection.}{1em}{}
\titleformat{\subsection}{\bfseries}{\thesubsection.}{1em}{}

\title{Gear List for Fate's Edge}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Gear List}

All gear items in Fate's Edge cost XP to acquire and require Downtime to obtain. Unless specified, gear is considered \textbf{Maintained} upon acquisition. Gear can become \textbf{Neglected} or \textbf{Compromised} through Story Beats or narrative events.

\subsection*{Weapons}

\begin{itemize}[leftmargin=*]
  \item \textbf{Dagger} (4 XP, 1 day): Melee 1 weapon. Can be concealed. Starts Controlled on stealth kills or parries.
  \item \textbf{Sword} (6 XP, 2 days): Melee 2 weapon. Standard melee combat effectiveness.
  \item \textbf{Bow} (6 XP, 3 days): Ranged 2 weapon. Effective at range; requires ammunition (arrows).
  \item \textbf{Crossbow} (8 XP, 4 days): Ranged 3 weapon. High damage; slower reload. Requires ammunition (bolts).
  \item \textbf{True-Steel Blade} (10 XP, 2 weeks): A masterwork weapon that never rusts or tarnishes. Does not require upkeep for maintenance purposes.
  \item \textbf{Hunter's Blade} (8 XP, 1 week): A well-balanced sword that sings slightly when blood is near. Grants start Controlled when making the first attack in an ambush or against unaware foes.
  \item \textbf{Throwing Needles (Set of 5)} (4 XP, 2 days): Small, concealable weapons. +1 die to Subterfuge when used for silent takedowns or distractions.
\end{itemize}

\subsection*{Armor}

\begin{itemize}[leftmargin=*]
  \item \textbf{Leather Jerkin} (4 XP, 2 days): +1 die to resist Fatigue from physical exertion. Provides minor protection.
  \item \textbf{Chainmail Hauberk} (8 XP, 1 week): +1 die to resist Harm from melee attacks. Heavy (counts as -1 die to stealth rolls).
  \item \textbf{Shield} (4 XP, 3 days): +1 die to defend against melee attacks. Can be used offensively for shove actions.
  \item \textbf{Shadowweave Cloak} (8 XP, 1 week): A cloak woven with dark threads. When stationary in dim light, wearer is treated as having +1 die to Stealth to hide. Does not require upkeep.
\end{itemize}

\subsection*{Tools \& Utility Gear}

\begin{itemize}[leftmargin=*]
  \item \textbf{Thief's Kit} (4 XP, 2 days): Includes lockpicks, rope, grappling hook, and small tools. +1 die to Subterfuge tasks involving locks, traps, or climbing.
  \item \textbf{Healer's Kit} (6 XP, 3 days): Bandages, herbs, and basic medical tools. +1 die to treat Harm or Fatigue. Can stabilize Critical injuries.
  \item \textbf{Scholar's Satchel} (4 XP, 2 days): Contains books, ink, parchment, and reference materials. +1 die to Lore or Research rolls.
  \item \textbf{Craftsman's Tools (Type)} (6 XP, 1 week): Specialized tools for a craft (e.g., blacksmith's forge, alchemist's lab). Required for complex crafting. +1 die to relevant Craft rolls.
  \item \textbf{Lantern \& Oil} (2 XP, 1 day): Provides light in dark areas. Negated by strong wind or water.
  \item \textbf{Climbing Gear} (4 XP, 2 days): Ropes, pitons, and harness. +1 die to climbing-related Athletics rolls.
  \item \textbf{Whisper Stone (Pair)} (6 XP, 1 week): Two stones that resonate when in contact with each other. Can transmit a single whispered sentence across any distance, once per day.
\end{itemize}

\subsection*{Mounts \& Vehicles}

\begin{itemize}[leftmargin=*]
  \item \textbf{Riding Horse} (8 XP, 1 week): Reliable mount. Provides fast travel overland. Requires feeding and care.
  \item \textbf{Cart \& Donkey} (6 XP, 1 week): Basic transport for goods. Slower than horse but carries more weight.
  \item \textbf{Swift Mare} (12 XP, 2 weeks): A spirited horse that moves with uncanny speed. When traveling overland, reduces travel clock segments by 1 (minimum 4). Does not require upkeep.
\end{itemize}

\subsection*{Magical Items}

\begin{itemize}[leftmargin=*]
  \item \textbf{Dream Journal} (6 XP, 1 week): A leather-bound book. Once per session, the owner can record a significant dream. The GM will use this to introduce a relevant omen or insight related to current challenges in a future session.
  \item \textbf{Compass Rose} (6 XP, 1 week): A compass that always points to a significant location chosen by the owner (e.g., home, a person, a place of power). Becomes unreliable if the target location is on another plane or is magically hidden.
  \item \textbf{Echoing Bell} (8 XP, 2 weeks): A small silver handbell. When rung once, its sound repeats exactly 10 minutes later. Useful for timekeeping or sending a delayed signal. The echo cannot be stopped once started.
  \item \textbf{Barkskin Cloak} (8 XP, 2 weeks): Made from the bark of an ancient tree. When worn, the wearer ignores the first instance of Minor Harm (-) each scene. Does not require upkeep.
  \item \textbf{Truthful Blade} (10 XP, 3 weeks): A simple dagger with a blade that gleams like starlight. When used to make a Sway roll based on honesty or to swear an oath, it grants start Controlled. If the wielder lies while holding it, it inflicts Harm 1 on them.
  \item \textbf{Lantern of the Lost Path} (10 XP, 2 weeks): A brass lantern that burns with a pale, cool flame. In wilderness areas, it prevents the party from getting lost due to natural causes (e.g., fog, darkness) and grants +1 die to Survival rolls for navigation. Does not require fuel.
  \item \textbf{Worry Stone} (4 XP, 3 days): A smooth, palm-sized stone. Once per session, when the bearer faces a significant social challenge (Sway, Sway, Deception), they can rub the stone to gain +1 die. The stone turns cold if the bearer is being deceived by someone else.
\end{itemize}

\subsection*{Specialty Gear}

\begin{itemize}[leftmargin=*]
  \item \textbf{Disguise Kit} (4 XP, 2 days): Wigs, makeup, and clothing for impersonation. +1 die to Deception rolls involving disguise.
  \item \textbf{Musical Instrument} (4 XP, 1 week): Required for Performance rolls. Quality instruments may grant +1 die.
  \item \textbf{Spyglass} (4 XP, 1 week): Allows long-distance observation. +1 die to Perception rolls involving distant objects.
  \item \textbf{Fine Clothes} (4 XP, 1 week): Proper attire for social situations. +1 die to Presence-based rolls in formal settings.
\end{itemize}

\end{document}
