
\Chapter{Legends}

\subsection{Saikou \& Amari (Tier IV)}

\subsubsection{Saikou Ira — "The Red Lamps See All"}
\textbf{XP Path:} Mastermind \quad
\textbf{Core:} A cursed strategist who reads positions—legal, social, criminal—the way others read faces.

\paragraph{Attributes \& Core Skills}
Wits 4, Presence 3, Spirit 3, Body 2.\\
Key Skills: \textbf{Insight 4}, Lore 4, Arcana 3, Command 3, Sway 3, Utaran Law \& Precedent 3.

\paragraph{Talent: Fogbound Read (6 XP)}
Once/scene, after you \emph{sell} an analysis (Method–Risk–Signal) of the current situation, roll \textbf{Wits + Insight}.\\
\emph{On success} choose two (on crit, choose all three):
\begin{itemize}
  \item \textbf{Exit Certainty:} ask the GM which single "safe" exit your foes think they have (truthful).
  \item \textbf{Bait Map:} declare one "obvious" reply your move appears to invite; it is a \emph{trap} (DV $-1$ vs that reply for your side this beat).
  \item \textbf{Noise Screen:} set Position to \emph{Controlled} for the next allied approach that exploits your read.
\end{itemize}
\emph{On failure:} GM banks +2 CP (\emph{you are being read}) and may raise DV +1 on the next opposed social/tactical action targeting your plan.

\paragraph{Talent: Oligarch's Cunning (7 XP)}
Once/session, instantly name a principal leverage (want/fear) in a negotiation; gain \emph{Controlled} for the ensuing parley.

\paragraph{Talent: Silver Tongue (4 XP)}
Gain +1 die when persuading or deceiving through speech.

\paragraph{Asset: Kahfagian Bloodline (Major, 12 XP)}
A standing social engine (charters, patronage, enemies).
\begin{itemize}
  \item \textbf{Invoke:} mark 1 \emph{Exposure} (Quiet$\to$Rumored$\to$Burned) \emph{or} spend 1 Boon to pull strings (counts as a small Diamond in ports/convoys).
  \item \textbf{Backfire:} on any GM CP spend in scenes you invoked it, the matriarch's \emph{price} arrives (debt, audit, rival).
\end{itemize}

\paragraph{Follower: Amari (Cap 4, 8 XP) — see below}

\subsubsection{Amari — The Loyal Sergeant}
\textbf{XP Path:} Solo \quad
Body 4, Spirit 3, Wits 2, Presence 2.\\
Key Skills: Melee 4, Athletics 3, Intimidation 3, Survival 2, Tactics 2.

\paragraph{Talents}
\textbf{Unstoppable Advance (8 XP):} When moving directly to remove a threat to Saikou, +1 effect and immunity to intimidation.\\
\textbf{Interpose (5 XP):} 1/scene, step in and take Harm meant for Saikou; roll your pool to reduce/negate.\\
\textbf{There Is No Puzzle, Only The Wall (4 XP):} Use Body+Melee to break non-combat obstacles (loudly).

\subsubsection{Table Use (tight loop)}
\begin{itemize}
  \item Saikou frames the scene with \textbf{Intricate} analysis (MRS) $\Rightarrow$ \emph{Fogbound Read} roll $\Rightarrow$ sets Position/DV tweaks and truth.
  \item Amari cashes the window with direct action; if plans wobble, he \emph{Interposes} or \emph{smashes through}.
  \item \textbf{GM CP spends} to hit the bond: tempt Amari with "clean" solutions that wreck stealth; tempt Saikou with Bloodline favors that cost Exposure.
\end{itemize}

\subsubsection{GM Levers (CP menu, tailored)}
\begin{itemize}
  \item \textbf{1 CP:} A rival "reads the read": +1 DV vs the declared bait this beat.
  \item \textbf{2 CP:} Bloodline push—gain immediate access, mark \emph{Exposure} on the Asset.
  \item \textbf{3 CP:} Red Lamp echo—convert the next mundane complication into an uncanny one bound to Saikou's insight.
  \item \textbf{4 CP:} Mother's writ closes a door you just opened; route changes or pay a Diamond-class concession.
\end{itemize}

\section{Riarhin — The Crimson Ghost (Epic Tier)}

\subsection{Role \& Doctrine}
\textbf{Callsign:} \emph{Red Mist}. \quad
\textbf{Type:} Covert ranger and "scout" (infiltration, deniable diplomacy, precision disruption). \\
\textbf{Doctrine:} \emph{"End the fight before it starts."} Shape the scene (Veil magic, oathcraft), secure  an exit, isolate the target, and exit cleanly—leaving a story that protects you.

\paragraph{Attributes \& Core Skills (SRD-aligned)}
Presence 4, Spirit 4, Wits 4, Body 2. \\
Skills: Stealth 4, Deception 4, Insight 4, Diplomacy 3, Melee 2, Survival 2, Lore 2, Arcana 3.

\paragraph{Art: Veil (Illusion-Weave)}
Treat \emph{Veil} as an \textbf{Art} in the Casting Loop (Channel with Wits+Arcana; Weave with Presence+Arcana). It persuades senses, mood, and attention; it never \emph{creates} mass or force.

\subsection{Signature Talents \& Methods}
\paragraph{Talent: Veil-Thread (6 XP)}
Once/scene, after an \emph{Intricate} description of the effect's \emph{seam} (what it shows, what it hides, who it anchors to), roll \textbf{Presence + Arcana}. \\
On success choose two (crit: all three):
\begin{itemize}
  \item \textbf{Soft Focus:} Downgrade one hostile's Position (Desperate$\to$Risky, Risky$\to$Controlled).
  \item \textbf{Borrowed Light:} Next social opener that leans into the mood gains +1 effect.
  \item \textbf{Kind Cut:} Convert up to 1 Harm from backlash into +1 Fatigue (you carry the price).
\end{itemize}
On failure: GM banks +2 CP; someone important notices the seam.

\paragraph{Talent: Name-Bound (7 XP)}
When you invoke a sworn name (oath, lineage, treaty) in good faith, gain \emph{Controlled} for the opening exchange and ask one narrow truth about the oath's current standing. Breaking or gaming it this scene gives the GM +2 CP.

\paragraph{Talent: Red Mist Entry (8 XP)}
On \emph{Approach} to a hostile site, if you establish an \emph{Intricate} entry plan with a prepped prop or costume \emph{on the table}, set your first action to \emph{Controlled} \textbf{and} treat one guard group as \emph{distracted} (−1 effect vs you) for one beat. \\
\emph{Note:} Re-rolling 1s does not remove CP; any new 1s add more CP.

\paragraph{Talent: Seam-Cut (5 XP)}
Once/scene, you can "cut" a rival illusion: roll \textbf{Wits + Insight}. On success, you either reveal its \emph{tell} (team gains +1 effect vs that illusion this scene) or \emph{shift} it one target away (it deceives someone else nearby).

\paragraph{Talent: Iron Stomach (3 XP)}
Immune to mundane poisons and spoiled food; halve Complications from toxic sources.

\subsection{Prestige: Mistweft Heir — Black Flag Methods (Epic)}
\paragraph{Prestige Ability: Bargain of the Red Mist (20 XP, 1/session)}
At a site of \emph{living memory} (ruin, oath-stone, barrow), you may ask a "why" that matters to the current scene and receive a true answer as an echo (perfect observation; no retcon). GM banks +2 CP. \\
\textbf{Hook — Balance Elsewhere:} If you use the truth to end a conflict cleanly (disarm, dissolve, de-escalate), the GM may dump CP to \emph{migrate the cost} to another bond/place you care about (Front tick).

\paragraph{Prestige Ability: Shadow Crown Method (18 XP, 1/campaign)}
For one scene, you may \textbf{name} a truce-line across a contested district. While active, violence across the line triggers immediate narrative calamity for the offender. Starts an 8-seg \emph{Coiled Blood} Front that passively ticks each session; when it blows, the rematch is worse—and your name is on it.

\subsection{Kit (Assets \& Diamonds)}
\paragraph{Major Asset: \emph{Remorse} (Relic Blade, 12 XP)}
Invoke: mark 1 \emph{Exposure} (your legend) or +1 Fatigue to assert \emph{dueling terms/truce-lines} for one beat. \\
Backfire: in presence of old empires or fate, GM may spend 2 CP to \emph{pull} you toward that gravity (detour clock).

\paragraph{On-Table Props (bring them physically if you can)}
\begin{itemize}
  \item \textbf{Way-cord (Valewood \SuitDiamond{})}: mark a return seam; +1 effect to exfil through it, but the corridor remembers.
  \item \textbf{Ward-salt / Lantern Writ (Mistlands \SuitDiamond{})}: safe pocket vs wraiths; chimes codify "no search" zones briefly.
  \item \textbf{Pilot Token / Harbor-Green Chit (Zakov \SuitDiamond{})}: priority lane or boom pass on the tide.
  \item \textbf{Foedus Seal (Vilikari \SuitDiamond{}) / Paiza (Ykrul \SuitDiamond{})}: frontier safe-conduct; offends the other law on CP spend.
\end{itemize}

\subsection{Associates (Followers, SRD rules)}
\begin{itemize}
  \item \textbf{Path-Warden (Cap 4, 8 XP)} — Survival/Stealth specialist; \emph{Assist} up to +3 dice (min(C, helper's Skill)).
  \item \textbf{Bell-Warden (Cap 3, 6 XP)} — Ward-salt, bell codes; converts one Mistlands \SuitClub{} to "warning" once/leg.
  \item \textbf{Pilot-Matron (Cap 3, 6 XP)} — Zakov lanes; once/leg negate a boom delay if given a \emph{Pilot Token}.
  \item \textbf{Sapper (Cap 3, 6 XP)} — Craft/Stealth; place non-lethal stoppers; can "cut noise" for an exfil beat.
\end{itemize}
\emph{Initiative Action:} 1 per scene party-wide; costs Exposure +1 or Harm 1 (SRD).

\subsubsection{Table Use (tight loop)}
\begin{enumerate}
  \item \textbf{Stage the Seam (Approach).} Intricate description (who sees what, what the seam hides) → Veil-Thread roll.
  \item \textbf{Anchor the Exit.} Place a Way-cord (Valewood) or Lantern Writ (Mistlands) \SuitDiamond{} face-up on table.
  \item \textbf{Bound the Contest.} If parley/duel is possible, draw \emph{Remorse} and speak the principle; lock escalation.
  \item \textbf{Cut Noise, Cut Tails.} Use Sapper/Pilot to keep clocks small; any 9 from Theona/Valewood adds an omission—plan for it.
  \item \textbf{Leave a Story.} Name-Bound or a witness-proverb secures your legend as the cover.
\end{enumerate}

\subsection{High-CP Spend Menu (GM, Riarhin-facing)}
\begin{itemize}
  \item \textbf{1 CP — Tell Glint:} An elite notices a seam; next deceive \emph{DV +1}.
  \item \textbf{2 CP — Echo Price:} Convert one clean success into \emph{Success \& Cost}; bank a \emph{Reputation Debt} (2/6) in-region.
  \item \textbf{3 CP — Counter-Story:} An adversary plants an official version; start a 4–6 seg \emph{Inspection/Inquiry} clock.
  \item \textbf{4+ CP — Balance Elsewhere:} Resolve today's scene in your favor, but immediately tick a connected Front and mark \emph{Legacy} in the Campaign Ledger (permanent world mark).
\end{itemize}

\subsection{Mission Seeds (Worked Missions with Deck Hooks)}
\paragraph{1) \emph{Quiet Water, Loud Bell} (Mistlands $\to$ Thin Shore)}
\textbf{Leg 1 Seed:} Mistlands \SuitSpade{} (Bell-Line levee), \SuitHeart{} (Bell-warden), \SuitClub{} (wraith crossing), \SuitDiamond{} (Ward-salt). \\
\textbf{Objective:} Ghost an Aeler audit on a friendly caravan. \textbf{Play:} Veil-Thread to look "already searched"; Ward-salt to hold line; exfil via Thin Shore.

\paragraph{2) \emph{Black Ladder} (Zakov insertion)}
\textbf{Leg 2 Seed:} Zakov \SuitSpade{} (Lantern Ladder), \SuitHeart{} (Lampman), \SuitClub{} (Informant Flip), \SuitDiamond{} (Smuggler's Ladder Map). \\
\textbf{Objective:} Swap a ledger shard at the Black Bell Tower. \textbf{Play:} Red Mist Entry; Seam-Cut a rival illusion; exit under Pilot Token.

\paragraph{3) \emph{Ninth Arch} (Theona truce-cut)}
\textbf{Leg 3 Seed:} Theona \SuitSpade{} (Uncounted Bridge), \SuitHeart{} (Matron of Wells), Linn \SuitClub{} (fogfall), \SuitDiamond{} (Moot Token). \\
\textbf{Objective:} Impose a 24-hour raid-truce to move a witness. \textbf{Play:} Remorse stance + Name-Bound; any counted step breaks Position.

\subsection{Backlash Palette (Veil Magic)}
\begin{itemize}
  \item \textbf{Lingering Mood:} The scene keeps the false tone; allies misread cues; \emph{Hearts} (Consequences deck) show in spades.
  \item \textbf{Echo Follows:} A person begins seeing you where you aren't (stalker or "witness" appears).
  \item \textbf{Borrowed Name:} An oath binds too broadly; you owe courtesy to someone you didn't mean to include.

\subsection{Backlash Palette (Glamour)}
\begin{itemize}
  \item \textbf{Lingering Mood:} The scene keeps the false tone; allies misread cues; \emph{Hearts} (Consequences deck) show in spades.
  \item \textbf{Echo Follows:} A person begins seeing you where you aren’t (stalker or “witness” appears).
  \item \textbf{Borrowed Name:} An oath binds too broadly; you owe courtesy to someone you didn’t mean to include.
\end{itemize}

\subsection{Advancement Hooks}
\begin{itemize}
  \item \textbf{Upgrade: Red Mist Entry $\to$ Ghost Rail (10 XP).} On a crit, you may also “lock” a single patrol path for one beat (they loop).
  \item \textbf{Upgrade: Seam-Cut $\to$ Unseam (9 XP).} Once/arc, turn a hostile illusion into a \emph{cover} for your team for one scene.
  \item \textbf{Prestige Path: Shadow Crown Method II (20 XP).} Expand the truce-line to a whole quarter for one watch; immediate \emph{Coiled Blood} +2 ticks.
\end{itemize}
