\section{The Amaranthine Sea and Surrounding Regions}

\subsection{Utar (The Utaran Imperium \& Successor Northlands)}
\label{sec:utar}

Origins \& Rise: Western tribes rose under Ecktor Utar, later becoming a republic, then the Imperium under Markus Ecktorius Leptis. Expanded eastward subduing various peoples.

Peak Administrative Map: At its height, held provinces Ecktoria, Hyrogia, Acasia, Vhasia, Viterra, with colonies Thedona and Zakovia.

Principal Cities (peak era):
\begin{itemize}
    \item Ecktoria (province): Ecktoria (capital), Canatha, Avennio, Histria
    \item Hyrogia: Thypergos (Thepyrgos), Sefina, Žunica
    \item Acasia: Sabiro (way-hub to Silkstrand interior)
    \item Vhasia: Lence, Vhaston, Ballafield, Eastwell
    \item Viterra: Tarlington, Valora, Fairport (royal port)
\end{itemize}

Decline \& Fracture: External shocks, plagues, the Great Ecktoria Quake, and the Fourth Ykrul War led to demographic collapse. The Imperium splintered into successor polities.

Successor Provinces:
\begin{itemize}
    \item Ecktoria: Wealthy rump-state clinging to forms and forums
    \item Acasia: Failed province of petty kings and Rothari warlords; Silkstrand the cosmopolitan exception
    \item Vhasia: Frontier, proud and touchy
    \item Viterra: Orderly, ducal powerbase; Valora ascendant
    \item Thepyrgos: Hyrogian heart around the city of Thypergos
\end{itemize}

\subsection{Acasia — "Broken Marches"}
\label{sec:acasia}

East of Ecktoria, Acasia is what happens when imperial carpentry rots but the beams still hold some weight. The province collapsed into internecine wars ~200 years after Utar's fall; today it's a mosaic of warlords, river-kings, toll dukes, and Rothari satraps. One true city remains: Silkstrand, a hard-nosed, cosmopolitan port that belongs to no one and profits from everyone.

Geography \& Routes:
\begin{itemize}
    \item The Green Lowlands: rolling farms and mill-rivers; crowded with palisade towns and toll bridges
    \item The Red Uplands: iron and ochre hills; ruined watchtowers, quarry hamlets, charcoalers
    \item The Reedbelt: labyrinth of floodplains and cut-canals feeding Silkstrand; smugglers' country
    \item Roads: Imperial stone breaks into causeway, then cart track
\end{itemize}

Power Map:
\begin{itemize}
    \item House Rothari (de facto hegemony): Ruthless logistics, hostage-ties, rotating "iron masters"
    \item Petty Kings \& Toll Dukes: Control bridges, mills, and market rights
    \item Silkstrand (free city): Ruled by a Triune - Harbormaster, Factor-Provost, Captain of Keys
    \item Church of the Flame: Wealthy abbeys; preach pacifism, sell grain futures
\end{itemize}

\subsection{Ecktoria — "Marble \& Fire"}
\label{sec:ectkoria}

A wealthy relic state that lives on memory and marble. The Grand Arena still crowns champions of the north; law schools and rhetors survive alongside bathhouses and clientage. Seat of the Church of the Flame and its most complete libraries.

Key Features:
\begin{itemize}
    \item The Grand Arena: Crowns champions of the north
    \item Law schools and rhetors: Centers of learning
    \item Church of the Flame: Vast estates, rural abbeys, mission colonies
    \item Countryside: Patchwork of Utaran villas, Kahfagian trader-towns, and Ikari enclaves
\end{itemize}

\subsection{Vhasia — "The Fractured Sun"}
\label{sec:vhasia}

Largest of the northern conquests and the last to fall, Vhasia is now a kingdom-in-name whose crown sits in Lence while real leverage lives with the Five Ducal Houses that once made it strong.

Present Polity:
\begin{itemize}
    \item Crownlands (Lence + fertile ring): Administered by Fenwood's governors
    \item The Five Duchies: De facto sovereigns with their own levies, coin, and laws
    \item Cadet "Blood" Houses: Names like Bloodgood, Longblood advertise lineage
\end{itemize}

Key Cities:
\begin{itemize}
    \item Lence: Former capital with museums, archives, hero-statues
    \item Vhaston: Trade city north of Vinmar; fast markets, faster tongues
    \item Eastwell: Big village with city habits; graft, protection rackets
    \item Ballafield: Frontier-turned-city; champion games, rough justice
    \item Marcott: River-port twin to Fairport (Viterra)
\end{itemize}

\subsection{Viterra — "The Last Kingdom"}
\label{sec:viterra}

Eastern of Vhasia, Viterra sprawls from Belworth River floodplains and peat Fens up through wind-gnawed Highlands, orchard-thick Dales, and the tangled Riverlands.

Geography:
\begin{itemize}
    \item Highlands: Sheep, iron blooms, slate; hillforts and watch beacons
    \item Fens \& Riverlands: Barges, peat, eels, flax; causeways and toll-stiles
    \item The Dales: Hedgerows, cider, barley; musters quick to form and quicker to gossip
    \item Belworth Corridor: Grain arteries, ferry rights, disputed parish maps
\end{itemize}

Cities \& Seats:
\begin{itemize}
    \item Tarlington: Marsh-edged ducal seat and administrative heart
    \item Valora: Old royal city; law courts, archives, Hall of Dawning
    \item Fairport: Tide-bitten river port at the Belworth delta
\end{itemize}

\subsection{Thepyrgos — "City of a Thousand Stairs"}
\label{sec:thepyrgos}

Last light of Astroegro—marble porticos over buried fire. Perched on the western face of the Astroego peninsula, it looks out across the Amaranthine like a stubborn lighthouse.

Key Features:
\begin{itemize}
    \item Terraced hills stitched with stairs, rampways, and aqueducts
    \item Vine-draped stoae that double as classrooms
    \item Roof-tile seas broken by slender watch-towers ("pyrgoi")
    \item Largest high-elf community on the Amaranthine, living openly but aloof
\end{itemize}

Institutions:
\begin{itemize}
    \item The Great University of Thepyrgos: Federation of colleges
    \item The Scriptorium Quarter: Copy houses, binders, illuminators
    \item The High-Elf Lyceum of Aethra: Teaches elven poetics, memory-arts
    \item The Conservatory of Nine Voices: Bards, court-singers, historians
\end{itemize}

\subsection{Silkstrand — "City of Bridges \& Dyewater"}
\label{sec:silkstrand}

Acasi trade oligarchy, Vinmar's rival. Cosmopolitan and decadent; its families rule like merchant princes. Home of the famed Bravos—flashy duelists who bleed as quickly as they boast.

Key Features:
\begin{itemize}
    \item Cosmopolitan and decadent culture
    \item Families rule like merchant princes
    \item Maintains a city watch, private armies, and a small navy
    \item Home of the famed Bravos—flashy duelists
\end{itemize}

\subsection{Kahfagia — "The Empire of Wakes and Storm-Flags"}
\label{sec:kahfagia}

Oligarchs at the helm; kraken-priests in the deeps; trade sails where war-galleys prowl. West of the Shoreless Bay and astride the Titan's Throat, Kahfagia holds the western quarter of the Amaranthine like a clenched gauntlet.

Geography:
\begin{itemize}
    \item The Titan's Throat: Narrow, current-wracked strait
    \item Outer Isles: Fortified harbors, drydocks, ropewalks
    \item The Western Shelf: Rich fisheries and kelp beds
    \item Terutria \& Sister Isles: Old Oshiiran cantonments turned admiralty towns
\end{itemize}

Cities:
\begin{itemize}
    \item Kassamira: Capital-harbor and dockyard labyrinth
    \item Stormspire: Black-basalt citadel on a headland
    \item Sable Hook: Privateer market
    \item Harth Quays: Marine barracks, tidepool training grounds
    \item Tereuth: Old Oshiiran grid turned Kahfagian guild town
\end{itemize}

\subsection{Theona — "The Marsh Crown"}
\label{sec:theona}

Three green isles of yeomen and bog, pledged to Viterra… when it suits them. On the Dolmis Sea east of Thepyrgos lie Atreus, Enoris, and Theona (largest), together scarcely 450 square miles of pasture, peat, and pale marble.

Geography:
\begin{itemize}
    \item Theona (Greater Isle): Low drumlins, peat bogs, hedged fields, white-veined marble outcrops
    \item Atreus: Wind-gnawed pasture, eel-choked creeks, stout hamlets
    \item Enoris: Stony ridges, orchard pockets, quarries with precarious cliff roads
\end{itemize}

\subsection{The Mistlands — "Bells, Salt, and Breath"}
\label{sec:mistlands}

Bread-basket valleys wrapped in wandering fog, farmed by daylight and warded by dwarven law. North of the Aelerian Mountains and east of the Direwood lies a quilt of river flats, hedge-laced fields, and low drumlins.

Geography:
\begin{itemize}
    \item Mist Flats: Deep loam, water meadows, willow-coppice
    \item Drumlin Belt: Long, whale-backed hills set with watch cairns
    \item The Fen Verge: Marsh where Direwood runnels peter out
    \item Stone Roads: Ancient, slightly-crowned pavements the fog rarely crosses
\end{itemize}

\subsection{Valewood — "The Forest That Remembers You Wrong"}
\label{sec:valewood}

Stretching from the Abderridan Sea's east coast into the far steppes, the Valewood carpets the rolling bones of the Anmont Divide. Its borders are curiously constant—villages on the rim can map its line—but the interior churns.

Geography:
\begin{itemize}
    \item North Vales (Taiga Rim): spruce and larch, bog-lights, frost-glass streams
    \item South Vales (Hardwood Mantle): beech, oak, ironwood
    \item Anmont Spurs: ruined terraces and causeways from an elder elven dominion
    \item Fixed Girdle: hamlets and toll-stiles on the outer roads
    \item Moving Heart: interior geography "precesses"
\end{itemize}

\subsection{Ykrul — "Storm on the Steppe" (Orcs)}
\label{sec:ykrul}

Horse-smoke on the horizon; bargains today, vengeance tomorrow. The Ykrul are a confederating people of riders and raiders native to the Violet Steppes and the forest margins of the Ebonwood.

Key Features:
\begin{itemize}
    \item Range: Violet Steeps, Ebonwood fringes, Aelerian foothills
    \item Seasons: pasture, omens, and opportunity
    \item Society: House → Band → Host structure
    \item Warfare: sudden strike, dispersed pursuit, vanish before noon
\end{itemize}

\subsection{Zakov — "Salt \& Serpent"}
\label{sec:zakov}

Pirate and crime haven, Dolmis Sea island. Controlled by the Seven Guilds, each a faction of thieves, smugglers, or slavers.

Key Features:
\begin{itemize}
    \item Power Structure: Controlled by the Seven Guilds
    \item Trade: Contraband, piracy, indentured servants
    \item Leadership: The so-called "King of Zakov" is an honorific, rotated between guild leaders
    \item Character: Lawless yet ritualized—its chaos has rules, written in blood and coin
\end{itemize}

Ports \& Locations:
\begin{itemize}
    \item The Salt-cracked wharf: where the tide brings back what the law forgets
    \item Bone-yard beach: wrecks and corpses tangled in kelp and rope
    \item The Shallows: reef-ridden cut only the locals dare
    \item The Iron Pier: gangplank to the Red Wake, a corsair flagship
    \item The Anchorhead: a floating fortress of the Pirate Syndicate
    \item The Serpent's Spine: a reef pass that leads to the city's secret heart
\end{itemize}

\subsection{Ubral — "The Stone Between Spears"}
\label{sec:ubral}

Clans, cairns, and hill-forts; iron oaths in a land that moves like mist. North of Viterra and east of Vhasia, Ubral climbs from wind-scoured low hills into the broken skirts of the Aelerian Mountains.

Geography:
\begin{itemize}
    \item High Moors \& Ridges: Heather, sheep, game; lookouts on every tor
    \item Glens \& Burns: Pocket farms, small cattle, flax
    \item Ore Country: Hill dwarf workings for iron and copper
    \item Passes North \& East: Narrow saddles to the Mistlands and Aelerian footholds
\end{itemize}

Cities \& Seats:
\begin{itemize}
    \item Dun Caerloch: Largest hill-fort and market-ring
    \item Stonebridge: Lowland river crossing turned town
    \item Keld Tor: Pass-watch holdfast
    \item Ravenmere: Peat-cutters' village grown into a trade hub
    \item Khaz Vurim: Hill dwarf surface depots and gate to deeper halls
\end{itemize}

\subsection{Linn — "Skerries \& Storm-Oaths"}
\label{sec:linn}

Fjords, skerries, halls, mistland routes. Southernmost outpost of the Linnic tribes—raiders, fur-hunters, and riverfolk from the Yrolka basin and the Alberriden Sea.

Key Features:
\begin{itemize}
    \item Culture: Southernmost outpost of the Linnic tribes
    \item Claim to Fame: They boast to have invented the dragon boats the Ykrul use
    \item Character: Fierce, pragmatic, half-tribal, half-mercantile
\end{itemize}

\subsection{Aelinnel — "Stone, Bough, and Bright Things"}
\label{sec:aelinnel}

Stone spires, sea-rock, deep woods. Tide-rift steps cut into black rock; moonwell basins in granite bowls; dolmen stairs up ridges.

Key Features:
\begin{itemize}
    \item Geography: Coastal stone formations, deep woods, tide pools
    \item Magical Elements: Moonlight motifs, antlers in shadow, petals that cut
    \item Inhabitants: Goat-herd of the stone edges, charcoal-burner clans, hedge-witches
\end{itemize}

\subsection{Aelaerem — "Hearth \& Hollow"}
\label{sec:aelaerem}

Hearth and hollow across rolling downs. Willow ford with flat stones, cider-press barns, chalk sheep-downs with turf mazes.

Key Features:
\begin{itemize}
    \item Landscape: Lanes, orchards, mills, downs, barrows
    \item Inhabitants: Hedge-witch midwives, millers with watch-geese, orchard reeves
    \item Folklore: Red thread motifs, quiet bells, watch-geese, footsteps under the soil
\end{itemize}

\subsection{Aeler — "Crowns \& Under-Vaults"}
\label{sec:aeler}

Crowns under-vaults across the Aelerians. Vaultmouth Gate in hillslope, Crown-Crypt Porch, Under-Market, Smoke-Shaft Stair.

Key Features:
\begin{itemize}
    \item Geography: Underground vaults, descents, underways, crown seats
    \item Inhabitants: Lamplighter apprentices, Under-Masons, Vault Wardens
    \item Culture: Stone/breath motifs, keys click, bells answer, mountain listens
\end{itemize}

\subsection{Black Banners — "Condotta \& Crowns"}
\label{sec:blackbanners}

Condotta and crowns. Burned village with field hospital, frozen ford with bodies, war-camp latrine row.

Key Features:
\begin{itemize}
    \item Warfare: Mercenary culture, condotta contracts, battlefield politics
    \item Geography: Camps, battlefields, fortresses, war-roads
    \item Inhabitants: Young cadets, veteran sergeants, Ecktorian banner-captains
\end{itemize}

\subsection{Vilikari — "Laurels \& Longhouses"}
\label{sec:vilikari}

Laurels and longhouses. Longhouse Quarter in old forum, Milefort XVII, Stone Ford with plank-palings.

Key Features:
\begin{itemize}
    \item Culture: Federates, mixed courts, diaspora peoples
    \item Geography: March towns, villa-forts, old roads
    \item Politics: Two laws motif - wolf and eagle side by side
\end{itemize}

\subsection{The Wilds — "Roads, Ruins, and Weather"}
\label{sec:wilds}

Roads, ruins, and weather. Crossing points, lookout knolls, old road traces, shelter hollows.

Key Features:
\begin{itemize}
    \item Flexibility: Reskin palette for any biome
    \item Geography: Varied terrain elements - crossings, lookouts, ruins, shelters
    \item Inhabitants: Forager children, guides, warden patrols, caravan crews
\end{itemize}

\subsection{Tulkani — "Road-Kin of the Ember Line"}
\label{sec:tulkani}

Painted wagons, braided oaths; songs for the living, bargains with the dusk. Once rooted northwest of Dhahara, scattered by Kuvani expansion and turned the road into a homeland.

Key Features:
\begin{itemize}
    \item Culture: Traveling wagon circles ("rings"), trade in various goods and services
    \item Organization: Rings led by Elders (keepers of fire) and Callers (negotiators)
    \item Religion: Kuva of the Hearth-Road, The Family (Raven road), Dawn \& Dusk Songs
\end{itemize}

\subsection{Ikari — "First Plough, First Oath"}
\label{sec:ikari}

Native tillers and smiths of the north; farmers at hearth, raiders at edge. The "first natives" of the continent north of the Amrathian Sea.

Tribes:
\begin{itemize}
    \item Kreki (Fishermen): net-masters, cod-curers, gamblers
    \item Ilikri (Wrights): carpenters, house-rafters, boat-builders
    \item Smeinnoii (Smiths): expert iron- and bronze-workers
    \item Glaesnirni (Glassers): furnace-builders and glassblowers
    \item Bhaellkidri (Brewfolk): malt and mead makers
    \item Samuldi (Seafarers): whale-hunters, raiders, storm pilots
    \item Ondriti (Law-Keepers): keepers of song-codes and adjudicators
\end{itemize}

\subsection{The Treaty of Flintpike}
\label{sec:flintpike}

The pact that made a market of the north—and a warning of what breaks it. Signed a year after the Battle of Flintpike on a field hedged with the skulls of oathbreakers.

Key Terms:
\begin{itemize}
    \item Common Weight \& Coin: Minted faces may differ; weights do not
    \item Trade Tongue: The "Common"—a broad-market pidgin formalized for writs
    \item Free Pass of the Marked: Anointed merchants with treaty-tokens
    \item Dwarf Assize: Disputes over measure, mint, and market settled by dwarf judges
    \item War Limits: No razing of mills, bridges, or dikes without formal war-bells
\end{itemize}

\subsection{Midh Adkaz — "Where War Became a Market"}
\label{sec:midhadkaz}

Frontier camp turned crossroads city; oaths traded as often as grain. Born as a mustering ground for the Second Ykrul War, refused to die when the banners furled.

Key Features:
\begin{itemize}
    \item The Red Ditch: Old siege-cut now serving as drainage and boundary
    \item Stakefield: Former picket-lines; today a grand fairground
    \item Boar Gate \& River Gate: Heavy gates controlling steppe and river traffic
    \item The Six Hands: Rotating council of steppe envoy, Ikari elder, dwarf factors, etc.
\end{itemize}

\subsection{Haayr Peninsula — "Anvils Between Two Seas"}
\label{sec:haayr}

Mountain tongues and broken coasts; whoever holds the passes writes the taxes. Jutting between the Dolmis Sea (north) and the Amrathian Sea (south and west).

Geography:
\begin{itemize}
    \item Spine of Haayr: Parallel limestone ranges with goat paths, watch-fires
    \item Northern Shelf (Dolmis coast): Harbor-studded bays, fishing towns
    \item Southern Folds (Amrathian coast): Terraced olives, marble quarries
    \item Interior Basins: Karst valleys with sinkholes, hot springs, grain pockets
\end{itemize}

Cities:
\begin{itemize}
    \item Khar-Myra: Northern cape-city; shipwrights, salt guilds, stone arsenal
    \item Theressos: Inland basin capital built around sulfur springs; healers' colleges
    \item Aigos-on-the-Falls: Market-town where three caravan roads meet a cliff portage
    \item Vekara: Southern marble port with half-drowned colonnades
\end{itemize}

Infrastructure:
\begin{itemize}
    \item The Hook Road: Cliff-hugging north–south artery
    \item Pass of Ten Towers: Main east–west saddle with beacon towers
    \item Haayr Ferries: Short sea hops between riven capes
\end{itemize}

\subsection{Dhahara — "Monsoon of Empires"}
\label{sec:dhahara}

Oath and oasis, monsoon and marchland; where caravans meet fleets under a war-bright sky. Far southeast of Galanina, past Sidhi marches and salt flats.

Regions:
\begin{itemize}
    \item The Himdal Marches (north): Steppe-edged uplands and loess plains
    \item Red Deserts \& Wadis (central): Wind-cut mesas, salt pans, seasonal waterways
    \item Incense Belt (western interior): Frankincense and myrrh terraces
    \item Monsoon Coast (south): Mangrove creeks and spice towns
    \item Jade Oases (east interior): Old stone rings, whispering vaults
\end{itemize}

Cities:
\begin{itemize}
    \item Sarvash: Caravan capital at crossing of three desert roads
    \item Hamun Gate: Fortress town guarding main pass from Marches
    \item Kassamira: Kahfagian presidial port; spice auctions at dawn
    \item Thalara: Incense-rich hill city with steps of a thousand cisterns
    \item Khetra Basin: Ring-oasis market abutting taboo precinct
\end{itemize}

\subsection{Oshiira — "The Ledger Empire"}
\label{sec:oshiira}

The ledger empire. A confederation-turned-state wedged between Shoreless Bay and the Crimson Basin.

Geography:
\begin{itemize}
    \item Core: Irrigated belts and canal webs between Shoreless Bay and Azaghari
    \item Rim: Fortified waystations and market towns on escarpments
    \item Roads \& Rivers: Numbered canal reaches and caravan "Lines"
\end{itemize}

Government:
\begin{itemize}
    \item Senate (Mbari-style) \& Magistracies: Rotating high offices
    \item Prefects: Rim administrators balancing census, embargo lists
    \item First Measure: State standard for weights/volumes
\end{itemize}

Military Traditions:
\begin{itemize}
    \item The Spirit of the Long Sorrow: Austere tradition born near Crimson Basin
    \item Arms \& Kit: Guardless blades, plain staves, boar-spears
    \item Practice: Slow breathwork, spine-locking footwork, kill-calm
\end{itemize}

\subsection{Sekogo — "Where the Roads Meet the Tide"}
\label{sec:sekogo}

Where the roads meet the tide. Northwestern third of the Crimson Basin, north of the Enjwe.

Geography:
\begin{itemize}
    \item North Rim (Oshiba Belt): Gallery forests, dry-season grazing
    \item Mid-Basin: Wet evergreen forests cut by blackwater creeks
    \item South Line (Enjwe Reach): Floodplains and levee towns
    \item Westward Lagoons: Barrier islands and mangrove lagoons
\end{itemize}

Government:
\begin{itemize}
    \item Circle of Unkwa: Rotating council of grove-masters, river-wardens
    \item Quay Syndicates: Licensed dock cartels bound by oath
    \item Lagoon Wardens: River pilots and coast guards
\end{itemize}

Principal City - Mbaro Quays:
\begin{itemize}
    \item The Tide Ledger: Auction hall where caravan lots split into hull-shares
    \item Brasswater Row: Repair yards for river craft
    \item Spice Shade: Colonnades trading pepper, kola, indigo
\end{itemize}

\subsection{Taharka — "Monsoon Crown, Terrace Throne"}
\label{sec:taharka}

Monsoon crown, terrace throne. Highland engine of canals, coins, and convoys.

Geography:
\begin{itemize}
    \item Mkusaro Highlands: Snowmelt reservoirs, stair-terraces, stone spillways
    \item Eastern Lowlands: Grid of irrigated fields with lift canals
    \item Southern Pastures: Transhumant herders rotating along state-posted grass calendars
    \item Riverworks: Southern Khesai checked by gates and levees
\end{itemize}

Government:
\begin{itemize}
    \item The Canal Collegium: Surveyors, hydraulic masons, sluice-masters
    \item Siatwe Senate: Merchant-landholders and mine-charterers
    \item The Mint of Siatwe: Sets regional coin and purity standards
\end{itemize}

Capital - Siatwe:
\begin{itemize}
    \item Terrace Steps Market: Spiraling bazaar mapped by water lanes
    \item Meltwater Court: Open square for sluice contracts and tenders
    \item Sea-Gate Countinghouse: Schedules monsoon convoys
    \item The Assay Hall: Public touchstones; verdict drums announce purity
\end{itemize}

\subsection{Ameria — "Between Bay and Throat"}
\label{sec:ameria}

Between Bay and Throat. North of the Titan Mountains, between the Shoreless Bay and the Titan's Throat.

Geography:
\begin{itemize}
    \item Shoreless Bay Coast: Broad harbors, fog belts, sea-winds
    \item Throatward Ports: Fort-crowned capes watching the Titan's Throat
    \item Interior Steppe: Vineyards, olive terraces, dusty roads
\end{itemize}

Government:
\begin{itemize}
    \item Split Crown: Royal style endures but west is Kahfagian-administered
    \item East maintains Neutral Regency recognized by both giants
    \item Consular Row: Hyro, Utaran, Kreki, Sidhi chanceries
\end{itemize}

Principal Cities:
\begin{itemize}
    \item Khol-Amar (Shoreless Bay): Customs Steps, Exchange Alley, Arcade of Maps
    \item Cape Verdan (Throatward Port): Signal Bluffs, Pilot Brotherhood, Chain-House
\end{itemize}

\subsection{Ngomebe — "Stone Men of the Moving Cities"}
\label{sec:ngomebe}

Stone men of the moving cities. First ironworkers of Akilan; wall-builders whose cities walk the mountains.

Geography:
\begin{itemize}
    \item Mkusaro South: Iron- and copper-bearing ridges, terraced goatland
    \item Ekale Spurs: Slate and granite quarries, cold streams
    \item Valley Roads: Stacked cairn-markers guiding caravans to guarded gates
\end{itemize}

Society:
\begin{itemize}
    \item Mason-Kin Houses: Hereditary wallers, quarry-masters, bridgewrights
    \item Iron Mothers: Matrilineal smelter guilds regulating ore allotments
    \item Gate-Voices: Masked stewards of each gate setting tolls
\end{itemize}

Principal Cities:
\begin{itemize}
    \item Khazembo (Ekale Spur): Stone Gate Row, Red Forge Court, Salt-Step Well
    \item Duma-Sete (Mkusaro Crest): "Walking" hill-city with folding walls
\end{itemize}

\subsection{Ashaan — "Gem of the Sea, Shadow on the River"}
\label{sec:ashaan}

Gem of the sea, shadow on the river. The fallen slaver-empire that once mastered the eastern Amaranthine rim.

Core Regions:
\begin{itemize}
    \item Core Seat: Ashaan/As'Ashaan, fortified river-island north of Khesai delta
    \item Green Ribbon / Great Desert: Fertile floodlands and canal belts
    \item Shadows of Ashaan: Galanina, Khasurah, northern Khesai villages
\end{itemize}

The Maeri - Three Sisters of Ashaan:
\begin{itemize}
    \item The Veiled Sister: Coins \& Winter; patroness of treasurers
    \item The Helmed Sister: Steel \& Sickness; patroness of warriors and assassins
    \item The Masked Sister: Red \& Black; patroness of art houses and revels
\end{itemize}

Power Structure:
\begin{itemize}
    \item The Esoti: Collegia of wizards and sorcerers interpreting Maeri's will
    \item The Black Hand: Fanatical network acting as Maeri's instrument
    \item Three Aspects: Veiled (accountants), Helmed (assassins), Masked (courtesans)
\end{itemize}

\subsection{Sihai — "The Central Kingdom, The Ordered Land"}
\label{sec:sihai}

The central kingdom, the ordered land. The ancient, immense, and rigidly hierarchical empire of the far east.

Geography:
\begin{itemize}
    \item The Heartland: Vast fertile plains of the Sihon River basin
    \item The Himadri Mountains: "Sky-Spine" separating from deserts and steppes
    \item The Hintara Ocean Coast: Long coastline with deep harbors
    \item The Altan Plains: Western grasslands, contested with Kuvani
\end{itemize}

Government:
\begin{itemize}
    \item The Son of Heaven: Emperor ruling with Mandate of Heaven
    \item The Bureaucracy: Meritocratic civil service via imperial examinations
    \item Doctrine: "The Five Relationships and Three Obediences"
\end{itemize}

Military:
\begin{itemize}
    \item The Imperial Army: Mass of disciplined infantry and elaborate logistics
    \item The Warrior Monks: Elite tradition of martial arts and philosophy
    \item The Navy: Controls Hintara Ocean and Nasan Sea
\end{itemize}

\subsection{Nihon — "The Isles of the Dawn Spirit"}
\label{sec:nihon}

The isles of the dawn spirit. A storm-wracked archipelago of fierce clans, divine spirits, and unparalleled artisans.

Geography:
\begin{itemize}
    \item The Isles: Chain of four main islands and countless smaller ones
    \item The Fire-Mountains: Several active volcanoes providing fertile soil
    \item The Inland Sea: Relatively calm body between main islands
\end{itemize}

Government:
\begin{itemize}
    \item The Heavenly Sovereign: Divine, hereditary emperor with ceremonial power
    \item The Shōgun: Theoretical military dictator ruling in emperor's name
    \item The Daimyō: Powerful regional lords ruling provinces with private armies
    \item The Samurai: Warrior aristocracy living by code of Bushidō
\end{itemize}

Military:
\begin{itemize}
    \item The Samurai: Elite core with mastery of bow, spear, and katana
    \item The Ashigaru: Commoner foot soldiers forming bulk of armies
    \item The Shinobi: Shadow-warriors specializing in espionage and assassination
\end{itemize}

\subsection{Ayokha — "The Monsoon Throne, The River of Heaven"}
\label{sec:ayokha}

The monsoon throne, the river of heaven. A sprawling, river-and-jungle fed kingdom famed for celestial temples.

Geography:
\begin{itemize}
    \item The Great River Basin: Lifeblood is the Sona River with annual floods
    \item The Jade Coast: Long coastline of mangrove swamps and natural harbors
    \item The Inner Jungle: Dense, mountainous rainforest isolating core
\end{itemize}

Government:
\begin{itemize}
    \item The Devaraja (God-King): Living embodiment of celestial spirit
    \item The Mandala System: Power radiates through vassal port-principalities
    \item The Celestial Bureaucracy: Class of priests managing rituals and legal codes
\end{itemize}

Military:
\begin{itemize}
    \item The Royal Guard: Elite warriors protecting Devaraja and capital
    \item War Elephants: Shock troops with armored beasts
    \item The Navy: True source of power with monsoon-riding junks
    \item Jungle Skirmishers: Light infantry from inland tribes
\end{itemize}

\subsection{Alberriden Sea — "Cold Mirror of the North"}
\label{sec:alberriden}

Cold mirror of the north. Brackish deeps, iron skies; river-teeth meet mountain-spines.

Geography:
\begin{itemize}
    \item The Yrolka Mouth ("River-Teeth"): Sprawling delta of mudflats and reed-beds
    \item The Deep Bowl: Central basins dropping black and cold
    \item Brack Marsh Rim: North and east shores dotted with peat flats
    \item Haravoa Shadow: Katabatic winds tumbling from mountains
\end{itemize}

Shores:
\begin{itemize}
    \item Mistland Coast (west): Low piers, mist-horns, root-cellar towns
    \item Valewood Edge (east): Few legal landings, standing-stone quays
    \item Northern Deltas: Eel weirs, peat hovels, winter ice-roads
    \item South Stone Harbors: Dwarf-cut steps and winches
\end{itemize}

\subsection{The Crimson Basin}
\label{sec:crimsonbasin}

Vast bowl-shaped rainforest drained by the Enjwe and web of blackwater tributaries. Control contested between wood elves and Oshiiran settlers.

Geography:
\begin{itemize}
    \item The Enjwe Trunk: Slow, splayed river with seasonal floods
    \item Blackwater Belts: Tannin-rich creeks flowing under closed canopy
    \item Rim Escarpments: Broken red cliffs where basin meets Ahmeric Plateau
    \item The Heartwood: Deep interior forest where canopy closes
\end{itemize}

Peoples:
\begin{itemize}
    \item Wood Elves (River Houses): Semi-nomadic clans tied to bends and groves
    \item Oshiiran Basinfolk: Confederated hamlets from Ahmeric Plateau migrants
    \item Treaty of Three Waters: Repeatedly renegotiated pact delimiting territories
\end{itemize}

Economy:
\begin{itemize}
    \item Forest Goods: Hardwoods, oils, fibers, dyes, spices, fish
    \item Farm-Forests: Mixed plots with yam, plantain, millet under shade trees
    \item Trade Corridors: Canoe-trains to Sekogo, portage to Ahmeric rim
\end{itemize}

\subsection{Linnstad}
\label{sec:linnstad}

Northern city-state, fur-trade hub. Southernmost outpost of the Linnic tribes—raiders, fur-hunters, and riverfolk.

Key Features:
\begin{itemize}
    \item Culture: Southernmost outpost of Linnic tribes
    \item Claim to Fame: Boast to have invented dragon boats used by Ykrul
    \item Character: Fierce, pragmatic, half-tribal, half-mercantile
\end{itemize}

\subsection{Rabelle}
\label{sec:rabelle}

Mountain city of ore and gem wealth. Rabellans are darker-skinned with striking red hair, claiming descent from half-giant scions of ancient Fire Giants.

Key Features:
\begin{itemize}
    \item Economy: Fine metalwork and trade with dwarves
    \item Position: Controls treacherous pass toward Mistlands
    \item Character: Blunt, proud, wealthy, fiercely protective of mountain identity
\end{itemize}

\subsection{Northpass}
\label{sec:northpass}

Frontier town, gateway to the Mistlands. Little more than caravan stop with notorious Ermine Inn.

Key Features:
\begin{itemize}
    \item Culture: Bare-bones outpost of vigilance
    \item Significance: Carefully monitored by dwarves
    \item Character: Rough, bare-bones, outpost of vigilance
\end{itemize}

\subsection{The Northern \& Eastern City-States}
\label{sec:citystates}

Beyond fractured provinces lie independent city-states, mountain strongholds, and coastal havens—bound by necessity, trade, and survival.

Shared Traits:
\begin{itemize}
    \item Independence: None bow easily to crowns
    \item Cultural Blend: Tribal, provincial, and foreign cultures collide
    \item Reputation: Dangerous, decadent, or dishonorable—but everyone trades
\end{itemize}

\subsection{Peoples of the Amaranthine}

\subsubsection{Wood Elves (Lethai-al)}
\label{sec:woodelves}

Mortal fey-descended humanoids of forests and jungles.

Key Features:
\begin{itemize}
    \item Distribution: Densest in northern temperate forests and Crimson Basin jungles
    \item Society: Motes of 20-120 kin led by keepers and callers
    \item Warfare: Avoid set-piece battles; ambush, misdirection, deniable pressure
    \item Magic: Low-signature, ecological—growth nudges, spoor-hiding, sense-bending fogs
\end{itemize}

\subsubsection{High Elves (Lethai-thora)}
\label{sec:highelves}

Mortal fey-descended humanoids of hidden upland vales and cloud-forests.

Key Features:
\begin{itemize}
    \item Distribution: Hidden citadels; largest enclave in Thepyrgos
    \item Society: Councils of Archivars, Pattern-wrights, Concord Keepers
    \item Warfare: Indirect power—treaty webs, sanctuaries, strategic illusions
    \item Magic: High-order geomancy, concordant wards, mnemonic magic
\end{itemize}

\subsubsection{Dwarves (Aeler)}
\label{sec:dwarves}

Vault builders, ledger keepers, soft power in hard rock.

Key Features:
\begin{itemize}
    \item Distribution: Mountain halls, hillholds, trade enclaves
    \item Economy: Guild networks standardizing weights, coin, stone-cuts
    \item Warfare: Fortify lines, starve sieges, collapse approaches
    \item Notable Orders: Spirit Shield Warriors, True Masons, Edgewalkers
\end{itemize}

\subsubsection{Halflings (Aelaerem)}
\label{sec:halflings}

People of the hearth. Green hills, deep larders, quiet courage.

Key Features:
\begin{itemize}
    \item Distribution: Amedell—a quilt of hedged paddocks and mill-creeks
    \item Society: Clans and hearths; matri/patrilineal with ledger inheritance
    \item Warfare: Hedgebows, bounders, burrow-works
    \item Culture: Plain speech, layered courtesy; hospitality with boundaries
\end{itemize}

\subsubsection{Gnomes (Aelinnel)}
\label{sec:gnomes}

People of sums. Hill-burrowed savants of glittering math, moss, and machinery.

Key Features:
\begin{itemize}
    \item Distribution: East of Viterra beyond Wrentfell River
    \item Society: Hearth-guilds, counting moots, dispute resolution by proof
    \item Magic: Low-glamour, high-precision ward-etching, probability slighting
    \item Craft: Lapidary, metalwork, terraced fungi, bee-keeping
\end{itemize}

\subsubsection{Ykrul (Orcs)}
\label{sec:ykrulpeoples}

Storm on the steppe. Confederating people of riders and raiders.

Key Features:
\begin{itemize}
    \item Range: Violet Steppes, Ebonwood fringes, Aelerian foothills
    \item Society: House → Band → Host structure
    \item Warfare: Sudden strike, dispersed pursuit, vanish before noon
    \item Culture: Blood tanistry, oath and omen, pasture diplomacy
\end{itemize}

\subsubsection{Ikari (Vilikari \& Ostrikari)}
\label{sec:ikaripeoples}

First plough, first oath. Native tillers and smiths of the north.

Key Features:
\begin{itemize}
    \item Distribution: Agrarian tribes of the northern continent
    \item Society: Chieftains, possible War-King in crisis, gender balance
    \item Warfare: Farmer-warriors with spear, round shield, axe
    \item Culture: Ancestral fires, sky and river rites, law by oath and ordeal
\end{itemize}

\subsection{Major Trade Routes and Waterways}

\subsubsection{Amaranthine Coastway}
\label{sec:amaranthinecoastway}

Kahfagia → Ecktoria → Acasia → Marcott (Vhasia) → Fairport (Viterra).

Key Features:
\begin{itemize}
    \item Draw Spade+Heart from port entered
    \item Draw Club from Wilds (gales, quarantine, harbor booms)
    \item Draw Diamond from port's deck or Kahfagia (permits, berths)
\end{itemize}

\subsubsection{Astroegro Straits}
\label{sec:astroegrostraits}

Thepyrgos controls the hinge between seas.

Key Features:
\begin{itemize}
    \item Draw Spade+Heart and Club from Thepyrgos
    \item Draw Diamond from Thepyrgos (harbor pass, crane allotment)
    \item Switch to Dolmis/Amaranthine modules once boom lowered
\end{itemize}

\subsubsection{Titan's Throat}
\label{sec:titansthroat}

Narrow strait controlled by Kahfagia between Amaranthine and outer seas.

Key Features:
\begin{itemize}
    \item Lighthouse chains, boom gates, submerged stakes
    \item Whoever times the tide owns the day
    \item Privateer diplomacy—today's sanctioned hunter tomorrow's apology
\end{itemize}

\subsection{Religious and Philosophical Traditions}

\subsubsection{Church of the Flame}
\label{sec:churchofflame}

Dominant faith of the Utaran successor states.

Key Features:
\begin{itemize}
    \item Doctrine: Adar, Odur, Akilesh as faces of Everflame vs Shadowflame
    \item Structure: Three High Luminaries, Honored Luminaries, College Radiant
    \item Ethic: "Make no war in the name of the Everflame" (observed selectively)
    \item Power: Vast estates, rural abbeys, mission colonies
\end{itemize}

\subsubsection{The Vigil}
\label{sec:thevigil}

Clandestine network bent on Utaran restoration.

Factions:
\begin{itemize}
    \item Steel Mantle: Covert, authoritarian, diplomatic ancestry
    \item Pale Sun: Bureaucratic apparatus; records and purse strings
    \item Crimson Talon: Paramilitary operations, deniable violence
\end{itemize}

\subsubsection{Faith of the Open Sky (Ykrul)}
\label{sec:ykrulfaith}

Nameless everywhere; shamans intercede with wind, storm, wolf, and ancestor.

Key Features:
\begin{itemize}
    \item Practices: Omen reading from sky, birds, horse-guts, fire-crack
    \item Taboos: Needless slaughter of herd-beasts, spilling guest-blood under host's roof
    \item Rituals: Sport and play as ritual bouts, dice, boast-feasts, board game koh'nreh
\end{itemize}

\subsubsection{Hearth and Hill Cults (Ubral)}
\label{sec:ubralfaith}

Ancestor fires, cairn offerings, oath taboos; breaking a sworn word is civil crime and spiritual stain.

Key Features:
\begin{itemize}
    \item Practices: Ancestor veneration, cairn maintenance, seasonal offerings
    \item Integration: Breaking sworn words creates both legal and spiritual consequences
    \item Coexistence: Everflame houses in southern glens, hermit traditions in ridges
\end{itemize}

\subsubsection{Water Rites (Various Coastal Peoples)}
\label{sec:waterrites}

Offerings at confluences, shrine maintenance, seasonal ceremonies.

Common Practices:
\begin{itemize}
    \item River pacts with salt, fish oil, and braided reeds
    \item Harbor shrines requiring libations from every launch
    \item Seasonal opening/closing of fishing waters
    \item Guest-right observances at water crossings
\end{itemize}

\subsection{Magical Traditions and Practices}

\subsubsection{The Eight Elements of Magic}
\label{sec:eightelements}

Fundamental system dividing magic into physical and metaphysical elements.

Physical Elements:
\begin{itemize}
    \item Earth: Solidity, stability, foundation
    \item Fire: Energy, transformation, destruction
    \item Wind: Movement, speed, freedom
    \item Water: Fluidity, healing, adaptability
\end{itemize}

Metaphysical Elements:
\begin{itemize}
    \item Fate: Destiny, inevitability, long-term consequences
    \item Life: Vitality, creation, spark of existence
    \item Fortune/Luck: Chance, unpredictability, probability
    \item Death/Dreams: End of things, transition, subconscious
\end{itemize}

\subsubsection{Arts of Magic}
\label{sec:magicarts}

Various methodologies for channeling magical potential.

Common Arts:
\begin{itemize}
    \item Alchemancy: Manipulation of chemical reactions and materials
    \item Herbomancy: Use of plants, medicinal and toxic
    \item Geomancy: Sensing or shaping the lay of the land
    \item Hydromancy: Predicting weather, finding water, calming currents
    \item Pyromancy: Tending flame, quick-starting fires, heat-hardening steel
    \item Illusiomancy: Veiling or crafting sensory deceptions
\end{itemize}

Forbidden Arts:
\begin{itemize}
    \item Thanatomancy: Full mastery over death and entropy
    \item Voidmancy: Drawing on spaces between worlds
    \item Domimancy: Subjugating minds, enslaving will
\end{itemize}

\subsubsection{High Elf Magical Philosophy}
\label{sec:highelfmagic}

Magic as conscious, responsible catastrophe management.

Core Principles:
\begin{itemize}
    \item The Law of Narrative Substance (The Echo): Everything possesses a metaphysical narrative
    \item The Law of Artful Manifestation (The Art): Framework for channeling potential
    \item The Law of Narrative Recoil (The Backlash): Magic changes problems rather than solving them
\end{itemize}

\subsection{Military Organizations and Warfare}

\subsubsection{Free Companies and Condotta}
\label{sec:freecompanies}

Mercenary units operating under contract between various powers.

Characteristics:
\begin{itemize}
    \item Dual loyalty to coin and current paymaster
    \item Professional standards maintained between contracts
    \item Condotta system allowing switching sides when advantageous
    \item Integration with local forces during extended campaigns
\end{itemize}

\subsubsection{Knights of the Dawn (Viterra)}
\label{sec:dawnknights}

Chivalric order sworn "to wake before the danger."

Key Features:
\begin{itemize}
    \item Dawn musters, road-keeping, hostage exchange
    \item Clean logistics and precise timing
    \item Less glitter, more grit than typical knightly orders
    \item Natural escorts for delicate parleys and logistical tasks
\end{itemize}

\subsubsection{Spirit Shield Warriors (Aeler)}
\label{sec:spiritshields}

Tattooed champions calling on ancestral wards to shelter allies.

Key Features:
\begin{itemize}
    \item Wandering arbiters of honor who claim tutelage from Hemrir
    \item Even hotheads yield to their judgments
    \item Revered within dwarven society
    \item Combines martial prowess with ancestral connection
\end{itemize}

\subsection{Economic Systems and Trade}

\subsubsection{The Way of Silk}
\label{sec:wayofsilk}

Major trade artery connecting western and eastern realms.

Key Points:
\begin{itemize}
    \item Western trunk: Silk, perfumed oils, dyes, glassware
    \item Central trunk: Barge-and-beast relay through canal systems
    \item Northern skirt: Fast "cloud caravans" of spare mounts
    \item Southern trunk: River transport with canal permits
\end{itemize}

\subsubsection{Coin and Currency Standards}
\label{sec:currency}

Various regional monetary systems with cross-border recognition.
 
Major Currencies:
\begin{itemize}
    \item Imperium Standards: Royals, Iron, Bits, Pennies with established conversions
    \item Dwarven Coinage: Silver, Copper, Dwar-Pennies with assay marks
    \item Thepyrgos Crowns: Silver coins trusted widely for weight and alloy
    \item Kahfagia Trident Coin: Maritime trade standard
\end{itemize}

\subsubsection{Trade Guilds and Merchant Organizations}
\label{sec:guilds}

Professional associations maintaining standards and protecting interests.

Notable Guilds:
\begin{itemize}
    \item Dockyard Guilds: Control ship construction and repair standards
    \item River Carter Syndics: Organize grain transportation
    \item Brewer and Smith Guilds: Maintain quality standards in Aeler holds
    \item Spice Syndicates: Control Fhara trade routes and oasis markets
\end{itemize}

\subsection{Cultural Practices and Social Customs}

\subsubsection{Guest-Right Traditions}
\label{sec:guestright}

Sacred custom observed across many cultures with regional variations.

Common Elements:
\begin{itemize}
    \item Three nights without question under host's roof
    \item Salt and bread ceremonies establishing protection
    \item Violation brings lasting shame and blood feud
    \item Transcends political boundaries and conflicts
\end{itemize}

\subsubsection{Oath-Binding Practices}
\label{sec:oaths}

Formal commitment ceremonies with spiritual and legal weight.

Types of Oaths:
\begin{itemize}
    \item Blood fines and wergild tariffs settling disputes
    \item Tree-oaths with binding at marked kapok or iroko
    \item Guest-right bonds creating temporary family ties
    \item Trade oaths sealed with knotted reed rings
\end{itemize}

\subsubsection{Seasonal Festivals and Calendar Observances}
\label{sec:festivals}

Regular celebrations marking agricultural, religious, or historical events.

Major Observances:
\begin{itemize}
    \item Millmoot (Acasia): Spring water-rights festival
    \item Chainnight (Acasia): Founders' walls celebration
    \item Dawning (Ecktoria): Spring equinox high holy day
    \item Monsoon Bells (Taharka): Marking seasonal wind changes
\end{itemize}

\subsection{Languages and Communication}

\subsubsection{Common Tongue}
\label{sec:common}

Dwarven-based creole language serving as trade lingua franca.

Development:
\begin{itemize}
    \item Originated as Dwarven creole called "Utarran"
    \item Dwarves generally refuse to learn other "Man Tongues"
    \item Supplanted older Utaro trade language after Treaty of Flintpike
    \item Regional dialects reflect local influences
\end{itemize}

\subsubsection{Regional Language Groups}
\label{sec:languagerelations}

Family groupings showing historical connections between peoples.

Major Groups:
\begin{itemize}
    \item Utaro and Hyro: Related language group
    \item Rabellan and Ubral: Related mountain languages
    \item Sidhi, Fhara, and Elvish: Related eastern tongues
    \item Viterran, Ikari, and Linnic: Northern language family
    \item Vhas and Acasi: Vulgar Utaro derivatives
\end{itemize}

\subsubsection{Literary and Scholarly Traditions}
\label{sec:scholarship}

Written works and educational practices preserving knowledge.

Centers of Learning:
\begin{itemize}
    \item The Great University of Thepyrgos: Federation of specialized colleges
    \item Aeler Vault Archives: Underground repositories of dwarven knowledge
    \item Valewood Grove Schools: Contextual learning in natural settings
    \item Oshiiran Canal Academies: Practical engineering and logistics training
\end{itemize}

\section{Conclusion}

The Amaranthine region represents a complex tapestry of cultures, each with distinct traditions, beliefs, and practices that have evolved over centuries of interaction, conflict, and cooperation. From the maritime empires of Kahfagia to the steppe confederations of the Ykrul, from the scholarly heights of Thepyrgos to the forest mysteries of Valewood, these lands offer endless opportunities for exploration and adventure.

Understanding the interconnected nature of these societies—their trade relationships, religious tensions, military traditions, and cultural exchanges—provides the foundation for rich storytelling and meaningful gameplay in the Fate's Edge setting.
