\documentclass[11pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{array}

\geometry{margin=1in}
\titleformat{\chapter}{\Huge\bfseries}{}{0em}{}
\titleformat{\section}{\Large\bfseries}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Fate's Edge: Magic Reference}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\newenvironment{spellentry}[3]{%
  \begin{center}
    \begin{tabular}{|p{5.5in}|}
      \hline
      \textbf{#1} \hfill \textbf{Art:} #2 \hfill \textbf{DV:} #3 \\
      \hline
}{%
      \hline
    \end{tabular}
  \end{center}
}

\begin{document}

\chapter{Magic in Fate's Edge}

\section{The Nature of Magic}

Magic in Fate's Edge is not a tool of convenience but a dangerous negotiation with the fabric of reality. It is powerful, flexible, and transformative—yet every attempt to shape it carries risk. The dice never merely ask ``does it work?'' but always whisper ``what is the cost?''

\subsection{Volatility by Design}
Magic is not fully understood, even by its most adept practitioners. Every working pushes against boundaries that resist being bent. This inherent instability means that magic is always a story moment, even when it succeeds.

\subsection{Risk Embodied}
Each spell generates Complication Points (CP). These points do not vanish; they manifest as Backlash—unpredictable consequences that ripple outward. The caster's burden is not just in channeling power, but in bearing the cost of that power.

\subsection{Thematic Consequence}
Backlash is not arbitrary; it aligns with the opposing or uncontrolled element of the Art invoked. Flame flares out of control, shadows linger too long, storms roll beyond command. The metaphysical elements have their own will.

\section{The Eight Elements of Magic}

The fundamental metaphysical system that underlies all magical practice in Fate's Edge divides magic into eight distinct elements, each with its own nature and resonance.

\subsection{Physical Elements}

\subsubsection{Earth}
\textbf{Nature:} Solidity, stability, foundation\\
\textbf{Domain:} Stone, soil, metal, structure\\
\textbf{Manifestation:} Shaping, sensing, moving earth and stone\\
\textbf{Backlash Theme:} Rigidity, collapse, awakening

\subsubsection{Fire}
\textbf{Nature:} Energy, transformation, destruction\\
\textbf{Domain:} Flame, heat, light, change\\
\textbf{Manifestation:} Ignition, heating, burning, purifying\\
\textbf{Backlash Theme:} Uncontrolled spread, scorching, consuming

\subsubsection{Wind}
\textbf{Nautre:} Movement, speed, freedom\\
\textbf{Domain:} Air, breath, sound, flight\\
\textbf{Manifestation:} Moving objects, creating drafts, amplifying sound\\
\textbf{Backlash Theme:} Whipping, dispersal, loss of control

\subsubsection{Water}
\textbf{Nature:} Fluidity, healing, adaptability\\
\textbf{Domain:} Liquids, tides, rain, flow\\
\textbf{Manifestation:} Calming, directing, purifying water\\
\textbf{Backlash Theme:} Stagnation, flooding, contamination

\subsection{Metaphysical Elements}

\subsubsection{Fate}
\textbf{Nature:} Destiny, inevitability, long-term consequences\\
\textbf{Domain:} Time, prophecy, causality\\
\textbf{Manifestation:} Seeing future threads, influencing destiny\\
\textbf{Backlash Theme:} Paradox, temporal loops, unintended consequences

\subsubsection{Life}
\textbf{Nature:} Vitality, creation, spark of existence\\
\textbf{Domain:} Health, growth, birth, death\\
\textbf{Manifestation:} Healing, growth, vitality enhancement\\
\textbf{Backlash Theme:} Overgrowth, decay, life drain

\subsubsection{Fortune/Luck}
\textbf{Nature:} Chance, unpredictability, probability\\
\textbf{Domain:} Randomness, luck, serendipity\\
\textbf{Manifestation:} Influencing chance, creating opportunities\\
\textbf{Backlash Theme:} Bad luck streak, probability collapse, ironic reversals

\subsubsection{Death/Dreams}
\textbf{Nature:} End of things, transition, subconscious\\
\textbf{Domain:} Sleep, death, spirits, the void\\
\textbf{Manifestation:} Communication with the dead, dream manipulation\\
\textbf{Backlash Theme:} Hauntings, nightmares, spiritual attention

\section{Magical Arts and Their Manifestations}

The Arts of Magic represent specific methodologies for channeling magical potential. Each Art is associated with one or more elements and has its own thematic expression and backlash patterns.

\subsection{Common Arts}

\subsubsection{Alchemancy}
\textbf{Elements:} Earth, Water, Fire\\
\textbf{Focus:} Manipulation of chemical reactions and materials\\
\textbf{Methods:} Transmutation, potion brewing, material refinement\\
\textbf{Backlash Theme:} Uncontrolled reactions, toxic fumes, material instability

\subsubsection{Herbomancy}
\textbf{Elements:} Earth, Life, Water\\
\textbf{Focus:} Use of plants, medicinal and toxic\\
\textbf{Methods:} Healing poultices, poison brewing, growth acceleration\\
\textbf{Backlash Theme:} Allergic reactions, plant overgrowth, seasonal confusion

\subsubsection{Geomancy}
\textbf{Elements:} Earth, Fate\\
\textbf{Focus:} Sensing or shaping the lay of the land\\
\textbf{Methods:} Earth sensing, structural reinforcement, seismic manipulation\\
\textbf{Backlash Theme:} Earthquakes, structural collapse, stone awakening

\subsubsection{Hydromancy}
\textbf{Elements:} Water, Life\\
\textbf{Focus:} Predicting weather, finding water, calming currents\\
\textbf{Methods:} Weather prediction, water purification, current control\\
\textbf{Backlash Theme:} Flooding, drought, water contamination

\subsubsection{Pyromancy}
\textbf{Elements:} Fire, Transformation\\
\textbf{Focus:} Tending flame, quick-starting fires, heat-hardening steel\\
\textbf{Methods:} Flame creation, heat manipulation, fire resistance\\
\textbf{Backlash Theme:} Uncontrolled fires, burns, heat exhaustion

\subsubsection{Illusiomancy}
\textbf{Elements:} Wind, Death/Dreams\\
\textbf{Focus:} Veiling or crafting sensory deceptions\\
\textbf{Methods:} Visual illusions, sound mimicry, sensory confusion\\
\textbf{Backlash Theme:} Illusion feedback, reality distortion, perceptual confusion

\subsection{Forbidden Arts}

\subsubsection{Thanatomancy}
\textbf{Elements:} Death/Dreams, Fate\\
\textbf{Focus:} Full mastery over death and entropy\\
\textbf{Methods:} Raising the dead, stopping hearts, accelerating decay\\
\textbf{Backlash Theme:} Death's attention, undead rebellion, life force drain

\subsubsection{Voidmancy}
\textbf{Elements:} Death/Dreams, Fate\\
\textbf{Focus:} Drawing on spaces between worlds\\
\textbf{Methods:} Dimensional rifts, pocket dimensions, banishment\\
\textbf{Backlash Theme:} Reality holes, interdimensional entities, spatial dislocation

\subsubsection{Domimancy}
\textbf{Elements:} Fortune/Luck, Fate\\
\textbf{Focus:} Subjugating minds, enslaving will\\
\textbf{Methods:} Mind control, compulsion, memory manipulation\\
\textbf{Backlash Theme:} Mental backlash, willpower drain, psychic parasites

\section{High Elf Magical Philosophy}

High Elves approach magic as conscious, responsible catastrophe management. Their philosophy is built on three core principles:

\subsection{The Law of Narrative Substance (The Echo)}
Everything possesses a metaphysical narrative—a story that defines its nature and purpose. Magic works by interacting with these narratives rather than simply manipulating physical properties.

\subsection{The Law of Artful Manifestation (The Art)}
Magic requires not just power, but artistry. The caster must craft their intention with precision, understanding both the desired outcome and the story they are trying to tell.

\subsection{The Law of Narrative Recoil (The Backlash)}
Magic does not solve problems—it changes them. Every spell creates new complications by altering the narrative fabric of reality. Wise casters prepare for these shifts.

\section{Casting Procedure}

All spellcasting follows a structured sequence called the Casting Loop, unfolding across two phases of play: gathering strength, then weaving it into form.

\subsection{Channel}
The caster focuses, rolling Wits + Arcana to gather Potential. Each success becomes fuel for shaping the spell. Each 1 adds Complication Points immediately.

\subsection{Weave}
On the following turn, the caster rolls Wits + (Art) to shape Potential into a defined effect. The Description Ladder applies: Basic/Detailed/Intricate descriptions reduce or redirect Complication Points.

\subsection{Backlash}
Complication Points spent by the GM manifest as uncontrolled consequences. These are thematic to the Art and scale with the number of points spent.

\section{Backlash Severity Table}

\begin{tabular}{|m{1in}|m{4in}|}
\hline
\textbf{CP Range} & \textbf{Effect} \\
\hline
1–2 CP & Minor nuisance or tell; short-lived cost, noise, or reveal. \\
\hline
3–4 CP & Noticeable setback: a real hazard, condition, or new pressure clock. \\
\hline
5+ CP & Major turn: scene shifts, a new foe/clock enters, or severe condition. \\
\hline
\end{tabular}

\section{Spell Creation Guidance}

\subsection{DV Guidance}
\begin{itemize}
    \item DV 1 = subtle sense or tiny edge
    \item DV 2 = scene-shaping buff/patch in a small area
    \item DV 3 = potent, loud, or multi-target edge
    \item Ace-level miracles should demand a price (Fatigue/Harm, costly component, or Obvious in a hostile place)
\end{itemize}

\subsection{Writing Spells}
Write the spell in 3 lines:
\begin{enumerate}
    \item Name \& Art: pick an evocative verb+noun
    \item Effect: one clear board change
    \item Backlash ladder: 1–2 CP = tell/minor cost; 3–4 CP = new pressure; 5+ CP = scene-altering twist
\end{enumerate}

\section{Pre-Made Spells}

\begin{spellentry}{Cinder-Fist}{Pyromancy}{2}
\textbf{Effect:} Your hand ignites; for up to 3 beats, unarmed actions gain +1 effect. On a strong hit vs flammables, you may Hazard –1 once. Requires a free hand; Obvious.\\
\textbf{Backlash:} 1 CP flame gutters after one use; 2 CP caster takes Fatigue 1 and scorches sleeve; 3 CP unintended item ignites (start a small Hazard clock); 4+ CP fire wreathes arm (Harm 2) and draws attention.
\end{spellentry}

\begin{spellentry}{Stone-Sense}{Geomancy}{1}
\textbf{Effect:} Sense through contiguous stone (30 ft); learn flaws/layout. Grant Controlled to a single move/breach using this read.\\
\textbf{Backlash:} 1 CP echo of ancient pain (distraction); 2 CP muddled by a competing presence; 3 CP partial petrification (–1 die physical for the next scene); 4+ CP something attuned in the stone is alerted to you.
\end{spellentry}

\begin{spellentry}{Still the Currents}{Hydromancy}{2}
\textbf{Effect:} Calm a 10 ft water square for 1 beat. Either start Controlled for one crossing or Hazard –1 in that zone. Sustaining costs 1 Fatigue/beat.\\
\textbf{Backlash:} 1 CP water becomes too mirror-flat (stealth tell); 2 CP chop intensifies outside the zone; 3 CP area stays eerily still until disturbed (curiosity magnet); 4+ CP water turns foul/stagnant, inviting pests.
\end{spellentry}

\begin{spellentry}{Cloak of Shadows}{Umbramancy}{2}
\textbf{Effect:} In dim or darker light, target starts Controlled vs sight-based detection; bright light ends the effect. Does not help vs sound/scent.\\
\textbf{Backlash:} 1 CP slight self-blindness (–1 die sight checks); 2 CP whispering shadows create a faint tell; 3 CP you borrow light from elsewhere, leaving a conspicuously bright patch; 4+ CP a shadow-being takes interest.
\end{spellentry}

\begin{spellentry}{Storm's Edge}{Stormcraft}{3}
\textbf{Effect:} For this scene, your ranged attacks gain +1 effect and may push targets back 1 zone. Requires open air; Obvious.\\
\textbf{Backlash:} 1 CP wind snatches small items; 2 CP lightning grounds nearby (Hazard +1); 3 CP gale forces repositioning (lose next action); 4+ CP storm lingers beyond control (start Weather(6) clock).
\end{spellentry}

\begin{spellentry}{Mend Flesh}{Vitalism}{3}
\textbf{Effect:} Choose one: clear 2 Fatigue or step down 1 Injury level (max to Moderate). Requires stillness (no sprint/fight beat).\\
\textbf{Backlash:} 1 CP patient must rest soon or gain Fatigue 1; 2 CP neglected minor injury festers; 3 CP clumsy overgrowth (–1 die related physical until treated); 4+ CP life drawn from surroundings (plants wither, small life dies).
\end{spellentry}

\begin{spellentry}{Whispering Wind}{Aeromancy}{2}
\textbf{Effect:} Send a whispered message up to 1 mile per Arcana. Message is clear but cannot carry complex information.\\
\textbf{Backlash:} 1 CP message is overheard by unintended recipient; 2 CP wind carries false messages from area; 3 CP your voice is carried away (cannot speak above whisper for scene); 4+ CP wind elemental takes interest in you.
\end{spellentry}

\begin{spellentry}{Bloom of Life}{Herbomancy}{3}
\textbf{Effect:} Cause plants in a 20 ft area to bloom or grow rapidly. Can create cover, entangle foes, or provide fresh food.\\
\textbf{Backlash:} 1 CP plants grow in unwanted directions; 2 CP allergic reactions affect allies; 3 CP seasonal confusion (plants behave as wrong season); 4+ CP awakened plant spirit becomes territorial.
\end{spellentry}

\section{Deck-Based Spell Seed Generator}

When improvising, draw 2–3 cards. Use the suit to pick an Art, rank to set DV and a scope tweak. Face cards add a quirk; Aces add a price.

\subsection{Suit Arts}
\begin{itemize}
    \item \heartsuit Vitalism / Hydromancy: life, vigor, calm, flow, restoration
    \item \clubsuit Geomancy / Discipline: stone, structure, weight, binding
    \item \diamondsuit Pyromancy / Transformation: heat, light, change, urgency
    \item \spadesuit Umbramancy / Veil: shadow, silence, misdirection, fear
\end{itemize}

\subsection{Rank → DV \& Scope}
\begin{itemize}
    \item 2–4: DV 1 (self or tiny area; 1 target; 1 beat sense/edge)
    \item 5–9: DV 2 (small zone ∼10 ft; team-sized buff; 1 rail tick)
    \item 10, J, Q, K: DV 3 (loud, group-facing, or multi-rail influence)
    \item Ace: DV 3 + price (component, Fatigue 1, Harm 1, or an Obvious flare)
\end{itemize}

\subsection{Face-card Quirks}
\begin{itemize}
    \item J Brief mobility or reach boost (leap, glide, slip)
    \item Q Social resonance (hush, awe, dread)
    \item K Durability/impact bump (+1 effect on one hard task)
\end{itemize}

\subsection{Backlash Prompts by Suit}
\begin{itemize}
    \item \heartsuit (Vitalism/Water) overgrowth, exhaustion, stagnation, sympathetic drain
    \item \clubsuit (Stone/Discipline) rigidity, slow, echoing calls, guardians notice
    \item \diamondsuit (Fire/Change) flare, scorch, noise, uncontrolled spread
    \item \spadesuit (Shadow/Veil) whispers, self-blindness, cold spots, entities take interest
\end{itemize}

\section{Play Examples}

\subsection{Example 1: Combat Casting}
\textbf{Situation:} Lyra the Pyromancer faces bandits in a narrow alley.

\textbf{Channel Phase:} Lyra rolls Wits + Arcana (5 dice): 8, 6, 4, 1, 1 = 3 successes, 2 CP\\
\textbf{GM spends 1 CP:} A nearby barrel starts to smolder (minor environmental hazard)

\textbf{Weave Phase:} Next round, Lyra Weaves Cinder-Fist with Detailed description (re-rolls 1s): Wits + Pyromancy (5 dice): 9, 7, 5, 3, 6 = 4 successes\\
\textbf{Effect:} Her flaming fist grants +1 effect on melee attacks for 3 beats

\textbf{GM spends remaining 1 CP:} The smoldering barrel bursts into flame, creating a small Fire(4) clock as a hazard

\subsection{Example 2: Utility Casting}
\textbf{Situation:} Theron the Geomancer needs to sense what lies beneath a suspicious floor.

\textbf{Channel Phase:} Theron rolls Wits + Arcana (4 dice): 10, 6, 1, 1 = 2 successes, 2 CP\\
\textbf{GM spends 2 CP:} The stone whispers of ancient pain, imposing a brief distraction (–1 die on next action)

\textbf{Weave Phase:} Theron Weaves Stone-Sense with Intricate description (re-rolls all 1s and adds flourish): Wits + Geomancy (4 dice): 7, 5, 4, 3 = 3 successes\\
\textbf{Effect:} He senses a hidden chamber beneath, granting Controlled position for accessing it

\subsection{Example 3: High-Stakes Ritual}
\textbf{Situation:} The party needs to heal a critically wounded ally during a siege.

\textbf{Channel Phase:} Elena rolls Wits + Arcana (6 dice): 9, 8, 6, 4, 1, 1 = 4 successes, 2 CP\\
\textbf{GM spends 1 CP:} The life force drawn causes nearby plants to wither

\textbf{Weave Phase:} Elena Weaves Mend Flesh with Detailed description: Wits + Vitalism (6 dice): 10, 7, 5, 3, 2, 1 = 4 successes, 1 CP\\
\textbf{Effect:} She steps down the ally's Severe injury to Moderate

\textbf{GM spends remaining CP:} The life drawn from surroundings creates an unsettling stillness—small creatures flee the area, and an ominous quiet falls over the battlefield

\section{Cultural Magical Traditions}

\subsection{High Elf Magic}
High Elves practice magic as conscious catastrophe management, focusing on long-term consequences and narrative resonance. Their spells often have subtle effects but profound implications.

\subsection{Wood Elf Magic}
Wood Elves specialize in low-signature, ecological magic that works in harmony with nature. Their spells include growth nudges, spoor-hiding, and sense-bending fogs.

\subsection{Dwarven Magic}
Dwarves focus on geomancy and stone-working, creating wards and structural enhancements. Their magic is practical and enduring.

\subsection{Gnome Magic}
Gnomes practice low-glamour, high-precision magic including ward-etching and probability slighting. Their spells are precise but often have unexpected side effects.

\end{document}
