\documentclass[11pt,twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathpazo} % Palatino font
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{titletoc}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{sectsty}
\usepackage{hyperref}

% Page setup
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[LO]{Fate's Edge Lore Primer}
\fancyhead[RE]{Player Guide}
\renewcommand{\headrulewidth}{0.4pt}

% Section formatting
\sectionfont{\color{black}\normalfont\fontsize{16}{19}\selectfont}
\subsectionfont{\color{black}\normalfont\fontsize{14}{17}\selectfont}
\subsubsectionfont{\color{black}\normalfont\fontsize{12}{15}\selectfont}

% Paragraph spacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

\title{Fate's Edge: A Player's Lore Primer}
\author{}
\date{}

\begin{document}

\maketitle
\thispagestyle{fancy}

\begin{center}
\textbf{What are you willing to risk, and what are you willing to pay, to reshape the world around you?}
\end{center}

\vspace{1em}

\tableofcontents

\newpage

\section{Core Concepts}

\subsection{Narrative Time}
Actions in Fate's Edge are framed by story weight, not strict chronology:
\begin{itemize}[leftmargin=*]
    \item \textbf{A Moment}: A heartbeat, a glance, a single strike.
    \item \textbf{Some Time}: A few minutes for skirmishes or negotiations.
    \item \textbf{Significant Time}: Hours for travel or rituals.
    \item \textbf{Days}: Large-scale endeavors like marches or recovery.
\end{itemize}

\subsection{Story Beats (SB)}
Every time you roll a 1, you generate a Story Beat. These aren't just penalties—they're narrative fuel for the GM:
\begin{itemize}[leftmargin=*]
    \item \textbf{Escalation}: Drawing more enemies or raising stakes.
    \item \textbf{Exhaustion}: Draining time, resources, or positioning.
    \item \textbf{Exposure}: Revealing hidden actions or alerting foes.
    \item \textbf{Collateral}: Harm spilling onto allies or surroundings.
\end{itemize}

\subsection{Affinity}
Your race or culture grants you a unique \textbf{Affinity}—a narrative edge that makes certain Arts, skills, or actions more reliable. This isn't just a mechanical bonus; it's a metaphysical bond with the world.

\subsection{Prestige Abilities}
High-level talents unlocked by mastering cultural arts or philosophies. These are narrative milestones as much as mechanical ones.

\subsection{On-Screen vs. Off-Screen}
\begin{itemize}[leftmargin=*]
    \item \textbf{On-Screen}: Companions, hirelings, or allies who stand beside you in danger. They add dice pools and flavor, but can falter or die.
    \item \textbf{Off-Screen}: Taverns, estates, titles, or networks. They never swing a blade in combat, but shape the story between sessions.
\end{itemize}

\section{The World of Fate's Edge}

\subsection{Geography}
\begin{itemize}[leftmargin=*]
    \item \textbf{Amaranthine Sea}: Western sea bordering Kahfagia and Ecktoria.
    \item \textbf{Dolmis Sea}: Inner eastern sea connecting several major powers.
    \item \textbf{Astroegro Straits}: Strategic waterway controlled by Thepyrgos.
    \item \textbf{Yloka Road}: Northern trade route through Linn territories.
\end{itemize}

\subsection{Major Regions}
\begin{multicols}{2}
\begin{itemize}[leftmargin=*]
    \item \textbf{Acasia}: Broken Marches—Cursed lawless territory.
    \item \textbf{Aelaerem}: Hearth \& Hollow—Halfling downs and orchards.
    \item \textbf{Aeler}: Crowns \& Under-Vaults—Dwarven mountain kingdoms.
    \item \textbf{Aelinnel}: Stone, Bough, Bright Things—Gnomish coasts and forests.
    \item \textbf{Black Banners}: Condotta \& Crowns—Mercenary warbands.
    \item \textbf{Ecktoria}: Marble \& Fire—Imperial forums and arenas.
    \item \textbf{Kahfagia}: Pilot's Mirror—Western maritime empire.
    \item \textbf{Linn}: Skerries \& Storm-Oaths—Norse island clans.
    \item \textbf{Mistlands}: Bells, Salt, Breath—Marshy borderlands.
    \item \textbf{Silkstrand}: City of Bridges \& Dyewater—Acasian trade hub.
    \item \textbf{Theona}: Three Greens, No Ninth—Island kingdom with taboos.
    \item \textbf{Thepyrgos}: City of a Thousand Stairs—Tower-city of scholars.
    \item \textbf{Ubral}: Stone Between Spears—Highland passes and clans.
    \item \textbf{Valewood}: Empire Under Leaves—Ancient fae ruins.
    \item \textbf{Vhasia}: The Fractured Sun—Broken kingdom of chivalry.
    \item \textbf{Vilikari}: Laurels \& Longhouses—Frontier federated states.
    \item \textbf{Viterra}: The Last Kingdom—Unified island realm.
    \item \textbf{Ykrul}: Wolf Standards, Winter Camps—Steppe orc clans.
    \item \textbf{Zakov}: Salt \& Serpent—Corsair archipelago.
\end{itemize}
\end{multicols}

\section{Magic System}

\subsection{The Eight Elements}
\begin{itemize}[leftmargin=*]
    \item \textbf{Physical}: Earth, Fire, Wind, Water
    \item \textbf{Metaphysical}: Fate, Life, Fortune, Death/Dreams
\end{itemize}

\subsection{Magical Arts}
\begin{itemize}[leftmargin=*]
    \item \textbf{Common Arts}: Alchemancy, Herbomancy, Geomancy, Hydromancy, Pyromancy, Illusiomancy
    \item \textbf{Forbidden Arts}: Thanatomancy, Voidmancy, Domimancy
\end{itemize}

\subsection{Casting Procedure}
\begin{enumerate}[leftmargin=*]
    \item \textbf{Channel}: Roll Wits + Arcana to gather Potential.
    \item \textbf{Weave}: Roll Wits + (Art) to shape the spell.
    \item \textbf{Backlash}: Story Beats manifest as thematic consequences.
\end{enumerate}

\section{Travel and Exploration}

\subsection{Regional Travel Themes}
Each region has unique mechanics and motifs:
\begin{itemize}[leftmargin=*]
    \item \textbf{Acasia}: Curse mechanics; every Ace adds lingering omens.
    \item \textbf{Aelaerem}: Red thread motifs; quiet bells and watch-geese.
    \item \textbf{Aeler}: Stone/breath motifs; keys click, bells answer.
    \item \textbf{Valewood}: Empire echoes; structures that phase in and out.
    \item \textbf{Theona}: "No Ninth" custom; omissions and taboos.
\end{itemize}

\subsection{Deck-Based Exploration}
Travel uses regional 52-card decks where:
\begin{itemize}[leftmargin=*]
    \item \textbf{Spades} = Places
    \item \textbf{Hearts} = Actors/Factions
    \item \textbf{Clubs} = Complications/Threats
    \item \textbf{Diamonds} = Rewards/Leverage
\end{itemize}

\section{Character Advancement}

\subsection{Player Archetypes}
\begin{itemize}[leftmargin=*]
    \item \textbf{The Solo}: Invests primarily in Attributes and Skills.
    \item \textbf{The Mixed Player}: Balances self-growth with assets.
    \item \textbf{The Mastermind}: Prioritizes followers and networks.
\end{itemize}

\subsection{Experience Points}
Used to improve:
\begin{itemize}[leftmargin=*]
    \item Attributes: New rating × 3 XP
    \item Skills: New level × 2 XP
    \item Followers: Cap² XP
    \item Off-Screen Assets: Minor (4), Standard (8), Major (12) XP
\end{itemize}

\section{Combat and Conflict}

\subsection{Position States}
\begin{itemize}[leftmargin=*]
    \item \textbf{Controlled}: Advantageous position, minor consequences.
    \item \textbf{Risky}: Even odds, moderate consequences.
    \item \textbf{Desperate}: Disadvantaged, severe consequences.
\end{itemize}

\subsection{Harm Levels}
\begin{itemize}[leftmargin=*]
    \item \textbf{Minor (-)}: 1 SB on next 2 rolls
    \item \textbf{Moderate (=)}: 1 SB on next roll, -1 die
    \item \textbf{Severe (>) } 2 SB on next roll, -2 dice
    \item \textbf{Critical (†)}: 3 SB on next roll, out of action
\end{itemize}

\section{Resources and Management}

\subsection{Supply Clock}
Shared party condition representing food, water, and gear:
\begin{itemize}[leftmargin=*]
    \item \textbf{Full Supply (0)}: Well-equipped
    \item \textbf{Low Supply (2)}: Minor complications
    \item \textbf{Dangerously Low (3)}: Each character gains Fatigue
    \item \textbf{Out of Supply (4)}: Severe penalties
\end{itemize}

\subsection{Fatigue}
Represents exhaustion, hunger, and strain:
\begin{itemize}[leftmargin=*]
    \item \textbf{1 Level}: Re-roll one success on next roll
    \item \textbf{2 Levels}: Re-roll one success on each roll
    \item \textbf{3 Levels}: Re-roll two successes on each roll
    \item \textbf{4 Levels}: Collapse/KO until treated
\end{itemize}

\end{document}
