\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{multicol}

\pagestyle{fancy}
\fancyhf{}
\rhead{Fate's Edge Demo}
\lhead{Shop Demo Game}
\rfoot{\thepage}

\titleformat{\section}{\large\bfseries\color{blue}}{\thesection}{1em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}

\newcommand{\dice}[1]{\texttt{#1}}
\newcommand{\dv}[1]{\textbf{DV #1}}
\newcommand{\cp}[1]{\textbf{CP: #1}}

\begin{document}

\begin{center}
{\LARGE \textbf{FATE'S EDGE SHOP DEMO}} \\
{\large \textbf{"The Broken Bridge" - 45-Minute Quickstart}}
\end{center}

\section*{OVERVIEW}

\textbf{Duration:} 45 minutes \\
\textbf{Players:} 1-6 (3 pre-gen characters) \\
\textbf{Theme:} Consequences, moral choices, world reactivity

\section{SETUP (5 minutes)}

\subsection*{Premise}
The PCs are traveling merchants who've reached a broken bridge in Acasia. They need to cross to reach the market town of Silkstrand, but the bridge is damaged and a local "tax collector" demands payment.

\subsection*{Pre-Generated Characters}
\begin{enumerate}
\item \textbf{Renn the Scout} - Body 3, Wits 2, Melee 2, Stealth 2
\item \textbf{Mira the Trader} - Presence 3, Wits 2, Sway 3, Deception 2  
\item \textbf{Thane the Guard} - Body 3, Spirit 2, Melee 3, Command 2
\end{enumerate}

\subsection*{GM Preparation}
\begin{itemize}
\item Standard deck of playing cards
\item Tokens for Supply/Fatigue tracking
\item Simple map of bridge area
\item Character sheets (handout)
\end{itemize}

\section{SCENE 1: THE BROKEN BRIDGE (10 minutes)}

\subsection*{Setup}
Bridge partially collapsed, rough-looking bandits collect "tolls"

\subsection*{GM Framing}
\begin{quote}
"The ancient stone bridge spans a deep gorge ahead, but the center span has collapsed. Makeshift planks provide a treacherous crossing. Three rough-looking figures in leather armor stand at the near end, crossbows ready."
\end{quote}

\subsection*{Key NPCs}
\begin{itemize}
\item \textbf{Garrick "Tax Collector"} - Thug leader (Body 2, Wits 2, Melee 2)
\item \textbf{Two Bandit Guards} - Basic threats (Body 2, Melee 1)
\end{itemize}

\subsection*{Player Options}
\begin{enumerate}
\item \textbf{Negotiate/Pay} - Mira tries diplomacy/deception
\item \textbf{Sneak Across} - Renn attempts stealth crossing  
\item \textbf{Fight} - Thane leads direct assault
\item \textbf{Find Another Way} - Wits + Survival to locate alternate path
\end{enumerate}

\subsection*{Sample Resolution}
\begin{itemize}
\item \textbf{Sway}: Presence + Sway vs \dv{2} (Garrick wants easy coin)
\item \textbf{Stealth}: Wits + Stealth vs \dv{3} (dangerous bridge + guards watching)
\item \textbf{Fight}: Body + Melee vs \dv{2} (bandits are poorly trained)
\item \textbf{Survival}: Wits + Survival vs \dv{2} (old roadbed still visible)
\end{itemize}

\subsection*{Complications to Introduce}
\begin{itemize}
\item Bridge is unstable (\cp{Hazard risk})
\item More bandits hidden (\cp{reinforcements})
\item Local villagers watching (\cp{reputation consequences})
\end{itemize}

\section{SCENE 2: THE HIDDEN THREAT (10 minutes)}

\subsection*{Reveal}
The bandits are actually protecting refugees from a recent curse incident in Silkstrand. The "toll" money feeds displaced families.

\subsection*{GM Framing}
\begin{quote}
"As you approach the bridge, you notice makeshift camps in the gorge below - dozens of frightened families. One desperate mother whispers that something terrible happened in Silkstrand - people disappearing, strange echoes..."
\end{quote}

\subsection*{New Challenges}
\begin{itemize}
\item \textbf{Moral Dilemma}: Take money from refugees or find another way?
\item \textbf{Investigation}: Wits + Insight to learn about the curse
\item \textbf{Social}: Presence + Sway with refugee leader
\end{itemize}

\subsection*{Quick Curse Element}
\begin{itemize}
\item Anyone who crosses without acknowledging the refugees experiences "echoes" (minor haunting)
\item Players can investigate or ignore
\end{itemize}

\subsection*{Sample Hooks}
\begin{itemize}
\item "My son walked the same path yesterday... now he only speaks in riddles"
\item "The bridge stones weep at night - we think they remember the old wars"
\end{itemize}

\section{SCENE 3: THE CHOICE (10 minutes)}

\subsection*{Climax Options}
Players must choose between:
\begin{enumerate}
\item \textbf{Force Crossing} - Fight bandits, gain supplies but create enemies
\item \textbf{Pay Fair Toll} - Support refugees, gain local allies  
\item \textbf{Investigate Curse} - Risk danger for greater reward/revelation
\item \textbf{Find Alternate Route} - Safe but time-consuming
\end{enumerate}

\subsection*{Resolution Matrix}
\begin{center}
\begin{tabularx}{\textwidth}{|l|l|l|}
\hline
\textbf{Choice} & \textbf{Outcome} & \textbf{Consequence} \\
\hline
Force & Quick success, Supply +1 & Refugee Relations -2 \\
\hline
Pay & Moderate success, Supply -1 & Refugee Relations +2 \\
\hline
Investigate & Risky (\dv{3}), +1 Boon & Curse knowledge \\
\hline
Alternate & Slow success, Fatigue +1 & Safe passage \\
\hline
\end{tabularx}
\end{center}

\subsection*{Combat Quick-Reference}
\begin{itemize}
\item Bandits: Harm =, generate 1 CP per hit
\item Bridge hazard: Falling = Harm >
\item Position matters: Flanking = +1 die, high ground = start Controlled
\end{itemize}

\section{WRAP-UP \& PITCH (5 minutes)}

\subsection*{Quick Resolution}
Whatever choice, players face consequences. Curse element hints at larger world. Refugees remember kindness/betrayal.

\subsection*{Sample Endings}
\begin{itemize}
\item \textbf{Force}: "The refugees curse your name as you leave their suffering behind"
\item \textbf{Pay}: "The grateful mother presses a small charm into your hand - 'For your mercy'"
\item \textbf{Investigate}: "You realize this curse is just one of many... someone needs to stop them all"
\item \textbf{Alternate}: "You cross safely, but wonder about the families left behind"
\end{itemize}

\subsection*{The Pitch}
\begin{quote}
"What you just experienced was Fate's Edge - a game where every choice matters. That curse? It's part of a region-wide mystery. Those refugees? They're connected to Silkstrand's dark history. And your character's choice? It just changed the fate of everyone in this broken land."
\end{quote}

\subsection*{Key Selling Points}
\begin{itemize}
\item \dice{⚡} \textbf{Fast Resolution}: Actions take seconds, not minutes
\item \dice{🎲} \textbf{Meaningful Choices}: Every decision has consequences  
\item \dice{🌍} \textbf{Living World}: NPCs react to player actions
\item \dice{⚖️} \textbf{Risk/Reward}: Success always costs something
\item \dice{📈} \textbf{Growth}: Characters become legends through choices
\end{itemize}

\subsection*{Quick Character Sheet Handout}
\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{ATTRIBUTES (1-5)} & \textbf{SKILLS (0-5)} \\
\hline
Body - Physical & Melee - Weapons \\
Wits - Mental & Ranged - Missiles \\  
Spirit - Will & Athletics - Movement \\
Presence - Social & Survival - Wilderness \\
& Stealth - Hiding \\
& Sway - Talking \\
& Deception - Lying \\
& Insight - Perception \\
& Command - Leadership \\
\hline
\end{tabular}
\end{center}

\subsection*{XP Investment Preview}
\begin{itemize}
\item Next session: Investigate Silkstrand curse deeper
\item Future: Gain magical abilities, command armies, reshape nations
\item Always: Your choices echo through the world
\end{itemize}

\section{SHOP OWNER TIPS}

\subsection*{Why This Works for Demos}
\begin{enumerate}
\item \textbf{Self-Contained}: Complete story in 45 minutes
\item \textbf{Scalable}: Works with 1-6 players  
\item \textbf{Flexible}: Can emphasize combat, social, or mystery elements
\item \textbf{Hooked}: Curse element teases larger setting
\item \textbf{Memorable}: Moral choice creates emotional investment
\end{enumerate}

\subsection*{What to Highlight}
\begin{itemize}
\item "Every roll changes the story - not just success/failure"
\item "Power always demands a price in Fate's Edge"
\item "You're not just playing a character - you're shaping a legend"
\end{itemize}

\subsection*{Common Questions Answered}
\begin{itemize}
\item \textbf{Complexity?} "Simpler than D&D, more meaningful than FATE"
\item \textbf{Session Length?} "2-3 hours for full sessions, this is the quick version"  
\item \textbf{Setting?} "Medieval fantasy with consequences and mystery"
\item \textbf{Magic?} "Dangerous, thematic, always risky"
\end{itemize}

\begin{center}
\textbf{This demo showcases Fate's Edge's core strengths while fitting any shop's time constraints!}
\end{center}

\end{document}
