\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{tcolorbox}

\geometry{margin=1in}
\titleformat{\section}{\Large\bfseries}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Fate's Edge: Expanded Talents}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\newtcolorbox{talentbox}[1]{
    enhanced,
    sharp corners,
    boxrule=0.6pt,
    colback=white,
    colframe=black,
    title=#1,
    fonttitle=\bfseries
}

\begin{document}

\begin{center}
{\Huge Fate's Edge: Expanded Talents Compendium}\\
\vspace{0.5cm}
{\large New Abilities for Character Advancement}
\end{center}

\section*{General Talents}

\begin{talentbox}{Battle Instincts (6 XP)}
Once per scene, re-roll a failed defense roll.
\end{talentbox}

\begin{talentbox}{Silver Tongue (4 XP)}
Gain +1 die when persuading or deceiving through speech.
\end{talentbox}

\begin{talentbox}{Iron Stomach (3 XP)}
Immune to mundane poisons and spoiled food; halve Complications from toxic sources.
\end{talentbox}

\begin{talentbox}{Exceptional Coordination (8 XP)}
One follower can provide +4 assist dice.
\end{talentbox}

\begin{talentbox}{Hardened Resolve (5 XP)}
Once per session, ignore the first CP generated against you from fear or intimidation.
\end{talentbox}

\begin{talentbox}{Keen Observer (4 XP)}
Gain +1 die on rolls to notice hidden details or read social situations.
\end{talentbox}

\begin{talentbox}{Lucky Break (3 XP)}
Once per session, convert one failed roll to a partial success by spending 1 Boon.
\end{talentbox}

\begin{talentbox}{Quick Reflexes (5 XP)}
Once per scene, take an additional minor action when initiative allows.
\end{talentbox}

\begin{talentbox}{Steady Hand (4 XP)}
Immune to penalties from environmental factors on ranged attacks (wind, darkness, etc.).
\end{talentbox}

\begin{talentbox}{Tireless March (6 XP)}
Ignore the first level of Fatigue gained from travel or sustained activity.
\end{talentbox}

\section*{Racial and Cultural Talents}

\begin{talentbox}{Stone-Sense (Dwarves, 5 XP)}
Detect flaws in stone or earth; gain +1 die on Engineering or Craft rolls underground.
\end{talentbox}

\begin{talentbox}{Backlash Soothing (Wood Elves, 6 XP)}
Once per session, reduce a magical Backlash Complication by 2 points when in natural terrain.
\end{talentbox}

\begin{talentbox}{Blood Memory (Ykrul, 5 XP)}
After a battle, meditate to gain one temporary Skill die reflecting a foe's tactics for the next scene.
\end{talentbox}

\begin{talentbox}{Hearth Wisdom (Halflings, 4 XP)}
Once per session, gain +2 dice on rolls involving domestic tasks, hospitality, or community knowledge.
\end{talentbox}

\begin{talentbox}{Storm's Child (Linn, 5 XP)}
Gain +1 die on all rolls involving maritime activities or weather prediction.
\end{talentbox}

\begin{talentbox}{Desert Walker (Ikari, 4 XP)}
Ignore penalties for desert or arid environment travel; +1 die on survival rolls in such terrain.
\end{talentbox}

\begin{talentbox}{Mountain Born (Rabellan, 6 XP)}
No penalty for high altitude or mountain terrain; +1 die on climbing and stonecraft.
\end{talentbox}

\begin{talentbox}{Steppe Whisperer (Tulkani, 5 XP)}
Communicate basic concepts with steppe animals; gain +1 die when tracking or surviving on plains.
\end{talentbox}

\begin{talentbox}{Shadow Step (Gnomes, 7 XP)}
Once per session, move unseen in dim light for one scene without rolling.
\end{talentbox}

\begin{talentbox}{High Speech (High Elves, 6 XP)}
Speak and understand any language for one scene; gain +2 dice on scholarly or magical research.
\end{talentbox}

\section*{Combat Talents}

\begin{talentbox}{Blade Dancer (8 XP)}
When fighting with light weapons, convert one CP into a counterattack that deals +1 effect.
\end{talentbox}

\begin{talentbox}{Shield Wall (7 XP)}
Once per battle, grant +1 die to defense rolls for all allies within close range for one scene.
\end{talentbox}

\begin{talentbox}{Precise Shot (6 XP)}
Ignore the first range penalty when making ranged attacks; +1 effect on called shots.
\end{talentbox}

\begin{talentbox}{Disarming Strike (7 XP)}
On a successful melee hit, you may attempt to disarm your opponent as a free action.
\end{talentbox}

\begin{talentbox}{Rapid Reload (5 XP)}
Reduce reload time for crossbows and firearms by one step; ignore penalties for hasty reloading.
\end{talentbox}

\begin{talentbox}{Combat Momentum (9 XP)}
After a successful attack, gain +1 die on your next action if it's also an attack.
\end{talentbox}

\begin{talentbox}{Defensive Stance (6 XP)}
Once per scene, convert a risky position to controlled for one defensive action.
\end{talentbox}

\begin{talentbox}{Flanking Master (8 XP)}
Gain +1 effect when attacking enemies that are flanked or surrounded by allies.
\end{talentbox}

\begin{talentbox}{Iron Skin (7 XP)}
Reduce all physical harm by one level (severe becomes moderate, etc.).
\end{talentbox}

\begin{talentbox}{Second Wind (9 XP)}
Once per battle, remove one level of Fatigue and clear one CP immediately after taking harm.
\end{talentbox}

\section*{Social Talents}

\begin{talentbox}{Charismatic Aura (8 XP)}
Gain +1 die on all Presence-based rolls when dealing with crowds or groups.
\end{talentbox}

\begin{talentbox}{Silver-Tongued Diplomat (7 XP)}
Once per session, re-roll a failed negotiation or diplomacy roll with a detailed description.
\end{talentbox}

\begin{talentbox}{Intimidating Presence (6 XP)}
Gain +1 die on rolls to intimidate or coerce; opponents may generate 1 CP from fear.
\end{talentbox}

\begin{talentbox}{Empathic Insight (7 XP)}
Read surface emotions and intentions of others; gain +1 die on Insight rolls.
\end{talentbox}

\begin{talentbox}{Cultural Savant (5 XP)}
Gain +1 die when interacting with cultures different from your own; learn basic phrases quickly.
\end{talentbox}

\begin{talentbox}{Liar's Mask (8 XP)}
Gain +2 dice on deception rolls; others must generate 2 CP to see through your lies.
\end{talentbox}

\begin{talentbox}{Inspiring Leader (9 XP)}
Once per session, allow all allies to re-roll one failed die during a group activity.
\end{talentbox}

\begin{talentbox}{Haggler's Wit (4 XP)}
Gain +1 die on all trading, bargaining, and negotiation rolls involving goods or services.
\end{talentbox}

\begin{talentbox}{Courtly Manners (6 XP)}
Navigate complex social hierarchies with ease; gain +1 die on etiquette and protocol rolls.
\end{talentbox}

\begin{talentbox}{People Person (5 XP)}
Remember faces and names easily; gain +1 die on rolls involving social connections or recognition.
\end{talentbox}

\section*{Exploration Talents}

\begin{talentbox}{Trailblazer (6 XP)}
Ignore difficult terrain penalties when leading a group; others gain +1 die on travel rolls.
\end{talentbox}

\begin{talentbox}{Night Vision (5 XP)}
Ignore darkness penalties for vision-based rolls; see twice as far in low-light conditions.
\end{talentbox}

\begin{talentbox}{Cartographer's Eye (7 XP)}
Create accurate maps from memory; gain +2 dice on navigation rolls in explored areas.
\end{talentbox}

\begin{talentbox}{Weather Sense (6 XP)}
Predict weather changes up to 24 hours in advance; gain +1 die on weather-related survival rolls.
\end{talentbox}

\begin{talentbox}{Trap Finder (8 XP)}
Notice traps and hazards automatically; gain +2 dice on rolls to disarm or avoid them.
\end{talentbox}

\begin{talentbox}{Climber's Grip (5 XP)}
Ignore penalties for climbing and vertical movement; +1 die on acrobatics and balance rolls.
\end{talentbox}

\begin{talentbox}{Swimmer's Grace (5 XP)}
Move at full speed underwater; ignore penalties for aquatic movement and combat.
\end{talentbox}

\begin{talentbox}{Tracker Supreme (9 XP)}
Follow any trail regardless of age or conditions; gain +2 dice on tracking rolls.
\end{talentbox}

\begin{talentbox}{Forager's Knowledge (4 XP)}
Identify edible plants, safe water, and useful materials in wilderness environments.
\end{talentbox}

\begin{talentbox}{Beast Whisperer (7 XP)}
Communicate basic concepts with animals; gain +1 die when handling or calming beasts.
\end{talentbox}

\section*{Lore and Knowledge Talents}

\begin{talentbox}{Lorekeeper (4 XP, High Elf only)}
Recall obscure history or magic without rolling once per session.
\end{talentbox}

\begin{talentbox}{Weave Anchor (8 XP, Wits 3+, Arcana 3)}
Reduce magical Backlash by 1 when casting Intricate spells.
\end{talentbox}

\begin{talentbox}{Echo-Walker's Step (20 XP, High Elf only; Wits 5, Arcana 4)}
Step briefly into the Ways Between; once per arc, turn a Complication into a boon.
\end{talentbox}

\begin{talentbox}{Blood Frenzy (4 XP, Body 2+)}
When reduced to half health, gain +1 die on melee rolls.
\end{talentbox}

\begin{talentbox}{Blood Memory (7 XP, Body 3+)}
After battle, gain 1 temporary Skill die reflecting a foe's tactics in the next scene.
\end{talentbox}

\begin{talentbox}{Warglord (18 XP, Body 5, Leadership 3)}
Rally scattered warbands; once per campaign, unify tribes under one banner.
\end{talentbox}

\begin{talentbox}{Arcane Savant (10 XP)}
Gain +1 die on all Arcana rolls; reduce spell casting time by one step.
\end{talentbox}

\begin{talentbox}{Forbidden Knowledge (12 XP)}
Access one Forbidden Art with reduced backlash (–1 CP on all backlash effects).
\end{talentbox}

\begin{talentbox}{Ritual Master (15 XP)}
Lead ritual castings with up to 5 assistants; reduce ritual time by half.
\end{talentbox}

\begin{talentbox}{Elemental Affinity (8 XP)}
Choose one element; gain +1 die on all spells of that element and reduce its backlash by 1.
\end{talentbox}

\section*{Regional Talents}

\begin{talentbox}{Silkstrand Duelist (9 XP)}
Gain +1 die in formal duels; Bravos recognize your skill and may offer favorable odds.
\end{talentbox}

\begin{talentbox}{Mistlands Navigator (8 XP)}
Navigate the fog-choked waterways of the Mistlands without penalty; gain +1 die on bell-line travel.
\end{talentbox}

\begin{talentbox}{Valewood Wayfarer (10 XP)}
Move through the shifting forests without becoming lost; gain +1 die on survival rolls in fae territory.
\end{talentbox}

\begin{talentbox}{Steppe Rider (7 XP)}
Gain +1 die when riding and +1 effect on mounted combat actions.
\end{talentbox}

\begin{talentbox}{Dwarven Hospitality (6 XP)}
Gain +1 die on all social rolls with dwarves; access to under-vault services at reduced cost.
\end{talentbox}

\begin{talentbox}{Kahfagian Pilot (9 XP)}
Gain +2 dice on all maritime navigation and pilotage rolls; lantern-law shifts favor you.
\end{talentbox}

\begin{talentbox}{Theona Taboo-Keeper (8 XP)}
Understand and navigate the "No Ninth" customs; gain +1 die on social rolls in Theona.
\end{talentbox}

\begin{talentbox}{Aeler Breath-Brother (7 XP)}
Gain +1 die on all rolls in dwarven holds; breath-measure allotments extended by 50\%.
\end{talentbox}

\begin{talentbox}{Linn Sea-Sense (6 XP)}
Predict sea conditions and hazards; gain +1 die on all maritime survival and navigation rolls.
\end{talentbox}

\begin{talentbox}{Ubral Hill-Fortune (5 XP)}
Gain +1 die on all rolls involving guest-right customs and hill-clan traditions.
\end{talentbox}

\section*{Craft and Trade Talents}

\begin{talentbox}{Master Artisan (12 XP)}
Create items of exceptional quality; gain +2 dice on all crafting rolls and reduce material costs by 25\%.
\end{talentbox}

\begin{talentbox}{Alchemical Genius (10 XP)}
Brew potions and create alchemical items with reduced material costs; gain +1 die on alchemy rolls.
\end{talentbox}

\begin{talentbox}{Smith's Forge (8 XP)}
Create weapons and armor of superior quality; gain +1 die on all metalworking rolls.
\end{talentbox}

\begin{talentbox}{Merchant's Eye (7 XP)}
Identify valuable items and fair prices automatically; gain +2 dice on appraisal rolls.
\end{talentbox}

\begin{talentbox}{Shipwright's Hand (9 XP)}
Build and repair vessels with exceptional quality; reduce construction time by 25\%.
\end{talentbox}

\begin{talentbox}{Tinker's Touch (6 XP)}
Repair broken items with improvised materials; gain +1 die on all repair rolls.
\end{talentbox}

\begin{talentbox}{Engineer's Mind (11 XP)}
Design and construct complex mechanical devices; gain +2 dice on engineering projects.
\end{talentbox}

\begin{talentbox}{Glassworker's Vision (8 XP)}
Create glass items of exceptional clarity and beauty; gain +1 die on all glassworking rolls.
\end{talentbox}

\begin{talentbox}{Leatherworker's Craft (5 XP)}
Create leather goods of superior quality and durability; reduce material waste by 50\%.
\end{talentbox}

\begin{talentbox}{Stonemason's Art (7 XP)}
Carve and shape stone with precision; gain +1 die on all stoneworking rolls.
\end{talentbox}

\section*{Stealth and Infiltration Talents}

\begin{talentbox}{Shadow Walker (10 XP)}
Move silently in any environment; gain +2 dice on all stealth rolls.
\end{talentbox}

\begin{talentbox}{Lockpick's Touch (6 XP)}
Open locks without tools using improvised methods; gain +1 die on all lockpicking rolls.
\end{talentbox}

\begin{talentbox}{Face in the Crowd (8 XP)}
Blend into urban environments; gain +1 die on all disguise and infiltration rolls in cities.
\end{talentbox}

\begin{talentbox}{Safe House Network (12 XP)}
Maintain a network of secure locations in major cities; gain +1 die on all urban escape rolls.
\end{talentbox}

\begin{talentbox}{Eavesdropper's Ear (5 XP)}
Hear conversations from great distances; gain +1 die on all listening rolls.
\end{talentbox}

\begin{talentbox}{Escape Artist (9 XP)}
Escape from bindings and restraints automatically; gain +2 dice on all escape rolls.
\end{talentbox}

\begin{talentbox}{False Identity (11 XP)}
Maintain multiple believable cover identities; others must generate 2 CP to see through your disguise.
\end{talentbox}

\begin{talentbox}{Urban Tracker (7 XP)}
Follow people through crowds and urban environments; gain +1 die on urban pursuit rolls.
\end{talentbox}

\begin{talentbox}{Infiltrator's Grace (8 XP)}
Bypass security systems and guards with ease; gain +1 die on all infiltration rolls.
\end{talentbox}

\begin{talentbox}{Dead Drop Master (10 XP)}
Establish and use covert communication systems; gain +2 dice on espionage-related rolls.
\end{talentbox}

\section*{Healing and Medicine Talents}

\begin{talentbox}{Battle Surgeon (12 XP)}
Perform emergency medical procedures in combat; heal injuries one level faster than normal.
\end{talentbox}

\begin{talentbox}{Herbalist's Knowledge (6 XP)}
Identify and prepare healing herbs; create basic remedies without rolling.
\end{talentbox}

\begin{talentbox}{Empathic Healing (9 XP)}
Sense the nature of injuries and ailments; gain +2 dice on all healing rolls.
\end{talentbox}

\begin{talentbox}{Preventive Medicine (7 XP)}
Diagnose potential health issues before they become serious; gain +1 die on medical prevention rolls.
\end{talentbox}

\begin{talentbox}{Surgical Precision (10 XP)}
Perform complex medical procedures with exceptional skill; reduce surgery time by 50\%.
\end{talentbox}

\begin{talentbox}{Plague Doctor (11 XP)}
Treat and prevent disease outbreaks; gain +2 dice on rolls involving epidemic control.
\end{talentbox}

\begin{talentbox}{Pain Management (8 XP)}
Reduce the effects of pain and discomfort; allies ignore first level of Fatigue from injury.
\end{talentbox}

\begin{talentbox}{Regenerative Touch (13 XP)}
Accelerate natural healing processes; patients recover twice as fast from rest.
\end{talentbox}

\begin{talentbox}{Antidote Brewer (9 XP)}
Create antidotes for poisons and toxins; gain +1 die on all antidote preparation rolls.
\end{talentbox}

\begin{talentbox}{Spiritual Healer (10 XP)}
Treat spiritual and psychological trauma; gain +1 die on rolls involving mental healing.
\end{talentbox}

\end{document}
