\section{Adventure Seeds}

\emph{Three scenarios drawn from Saikou Ira's unsolved cases. GMs are free to adapt them to their campaigns. Each has left a mark on me that I cannot exorcise. Perhaps you will succeed where I could not.}

\subsection{The Lamplighter's Echo}

\emph{Case No. 847-AE-03. Status: Unresolved.}

\textbf{Setup:} In the Aeler mining town of \emph{Vent-Deep}, the Echo of a lamplighter killed in a gas explosion has begun to manifest during the night shift. The Echo does not harm anyone---it simply repeats the lamplighter's final act: ringing a warning bell that no one hears. But the bell's phantom chime confuses the miners, causing them to lose count of their breaths. Two have already fallen into unseen shafts. The mine foreman, a pragmatic Iron Avenger named \emph{Harken Veth}, wants to blast the tunnel where the lamplighter died. The lamplighter's widow, \emph{Sanna}, begs the party to find another way.

\textbf{Saikou's Observation:} \emph{``The Echo is not a haunting. It is a \textbf{question}---the same question the lamplighter asked in his last moment: ''Did anyone hear me?`` The widow is the only one who did. But she was not in the tunnel. She was at home, kneading bread. She heard the bell in her chest, not her ears. That is the difference between witness and witness.''}

\textbf{Complications:}
\begin{itemize}
  \item The Echo is not the lamplighter's death---it is his \textbf{final attempt to ring the warning bell}. The bell itself was salvaged and now hangs in the town's chapel. But the chapel is built on a threshold that the Hollow visits every ninth day.
  \item The lamplighter's name has been forgotten. The widow is the only one who remembers, but she is old and her memory is frayed. She fears she will forget before the next manifestation.
  \item The foreman has already set the charges. He is not cruel---he is exhausted. Two more accidents will close the mine and cost the town its livelihood.
\end{itemize}

\textbf{Resolution Paths:}
\begin{enumerate}
  \item \textbf{Ring the Bell:} The party must retrieve the bell from the chapel on the night the Hollow walks (test Spirit+Lore to avoid its attention). Ring it during the Echo's next manifestation while the widow speaks her husband's name. The Echo will cease. \textit{Cost:} The widow forgets the lamplighter's face. She says it is a fair trade.
  \item \textbf{Blast the Tunnel:} The foreman detonates the charges. The Echo transforms into a Poltergeist (Class IV). It will haunt the mine until the party performs a full exorcism---or until it pulls the foreman into the same shaft where the lamplighter died. \textit{Cost:} One death, at minimum.
  \item \textbf{Redirect the Echo:} A risky third path. The party can ring a different bell at a different threshold, convincing the Echo that its duty is complete. Requires a \texttt{[WITNESS]} from the mine's entire shift crew (Presence+Command, DV 5). On success, the Echo fades. On failure, it shatters into multiple Echoes, each haunting a different tunnel.
\end{enumerate}

\textbf{Timers:}
\begin{itemize}
  \item \textbf{Foreman's Patience [4]} --- ticks each day the party delays. When filled, the foreman blasts the tunnel without them.
  \item \textbf{Echo's Desperation [6]} --- ticks each time the Echo is ignored, each time a miner is injured. When filled, the Echo becomes a Poltergeist.
  \item \textbf{Widow's Memory [4]} --- ticks each night she struggles to remember her husband's name. When filled, she forgets. The lamplighter's name is lost forever.
\end{itemize}

\subsection{The Ghoul of the Sewers}

\emph{Case No. 852-PS-09. Status: Closed (unsatisfactory).}

\textbf{Setup:} In Port Shed, beneath the spice warehouses, a series of missing persons reports have been linked to the sewers. The victims are always found later---alive, but \emph{changed}. They speak of a woman who begged them not to scream, who wept while she fed. The party is hired by a Sidhi dock-master named \emph{Old Kes}. He does not want the ghoul destroyed. He wants to know \emph{why} she is still hungry.

The ghoul is a young woman named \emph{Zel}. She was cursed after drinking from a well contaminated by a broken Veiled Sister's seal. She does not want to kill; she is starving. The curse has turned her into a Starved Anchor (Class II transition to Ghoul). She remembers her mother's face. She does not want to forget.

\textbf{Saikou's Observation:} \emph{``I have sat with ghouls who ate their own fingers in hunger. Zel was different. She did not eat. She \textbf{collected}---torn buttons, a child's shoe, a wedding ring. She was not feeding her body. She was feeding her \textbf{memory} of being human. That is worse. That is a debt she cannot pay because the currency is no longer minted.''}

\textbf{Complications:}
\begin{itemize}
  \item Zel's mother, \emph{Anja}, is still alive. She has been leaving bowls of animal blood at the sewer entrance for three months. She knows what her daughter has become. She does not care.
  \item Old Kes knows about Anja. He has been paying the mother to keep her quiet. He wants the party to \emph{deal with} the ghoul---quietly, permanently, without involving the Admiralty.
  \item A rival smuggler, \emph{Rowan the Eel}, has learned of the ghoul. He plans to capture her and sell her to an Ashaani flesh-mage as a ``living weapon.'' He has already hired a band of rat-catchers to corner her.
\end{itemize}

\textbf{Resolution Paths:}
\begin{enumerate}
  \item \textbf{Kill the Ghoul:} Straightforward, but morally ambiguous. Zel will not fight back---she will weep and ask for her mother. \textit{Cost:} Anja learns of the killing and places a \textbf{Blood Debt} on the party. Old Kes pays them, but they are haunted by her forgiveness.
  \item \textbf{Arrange a Feeding Schedule:} Negotiate with the dock-master, the mother, and a sympathetic butcher. Zel agrees to feed on animal blood only, under witness. Requires a \texttt{[PROMISE CLOCK] [4]}. If broken, Zel returns to the sewers---and this time, she will not weep. \textit{Cost:} Old Kes now owns a favor from the party. He will call it in.
  \item \textbf{Attempt a Cure:} Journey to the Well of Ichor (see \emph{Malachai, the Chained Angel}). The cure requires Zel to drink from the well and \emph{forget her hunger}---and her mother's face. Anja must give her permission. If she does, Zel becomes human again, but hollow. \textit{Cost:} Anja dies of grief within the year. The party gains a \textbf{String} from the Breath-less network for saving a soul, but also a \textbf{Debt} from the Veiled Sisters for ``interfering with a sealed curse.''
\end{enumerate}

\textbf{Timers:}
\begin{itemize}
  \item \textbf{Hunger [4]} --- ticks each day Zel does not feed. When filled, she attacks indiscriminately. The first target will be her mother.
  \item \textbf{Smuggler's Scheme [6]} --- ticks each time the party delays, each time Rowan gathers information. When filled, he captures Zel and transports her to Ashaan.
  \item \textbf{Mother's Patience [4]} --- ticks each time the party refuses to act. When filled, Anja goes into the sewers herself to bring her daughter home.
\end{itemize}

\subsection{The Possessed Censor's Daughter}

\emph{Case No. 859-EC-04. Status: Closed (with conditions).}

\textbf{Setup:} In Ecktoria, the Censor \emph{Fabius Valerius} is a man who weighs grain and judges contracts. He has never believed in anything except the ledger. His daughter, \emph{Livia}, fourteen years old, has been accused of witchcraft after her body began to speak in languages she never learned---old Utaran, Pereshi courtroom cant, and once, a dialect that made the Inquisition's translator weep.

The Everflame Inquisition has arrested her. They plan to burn her at the next Triumph. Fabius refuses to speak with the party---but his ledger has a hidden page, and the page leads to a name.

\textbf{Saikou's Observation:} \emph{``The demon's name was not a word. It was an \textbf{interest rate}---the compounding debt on a loan Fabius took to pay for his wife's medicine. The loan was sold to a collector, who sold it to a broker, who died. The interest compounded into a number that could never be paid. The demon was that number, given will. It had been living in his ledger for three years before it jumped to his daughter. The girl was not possessed. She was collateral.''}

\textbf{Complications:}
\begin{itemize}
  \item Fabius will not confess. He has spent his life enforcing contracts; he cannot admit that he broke one.
  \item The demon's true name is recorded in a forbidden text held in Fabius's private archive. The archive is sealed with a ward that only The Sentinel can breach. The ward is keyed to Fabius's blood.
  \item The Inquisition's schedule is fixed. The party has three days before the Triumph. The girl is held in a cell lined with silvered mirrors---the Inquisition's crude attempt to contain the demon. The mirrors are cracking.
  \item A faction within the Inquisition (the \emph{Chain-Lanterns of the Quiet Seal}) believes the girl may be innocent. They will help the party---but only if they promise to bring Fabius to justice afterward.
\end{itemize}

\textbf{Resolution Paths:}
\begin{enumerate}
  \item \textbf{Expose the Censor:} The party breaks into Fabius's archive, finds the forbidden text, and learns the demon's true name. They present the evidence to the Inquisition. Fabius is arrested; Livia is freed. The demon is exorcised by the Chain-Lanterns (with the party's help). \textit{Cost:} Fabius dies in a debtor's cell within the year. Livia never speaks again---not because she cannot, but because she has nothing left to say.
  \item \textbf{Secret Exorcism:} The party performs The Sentinel's Exorcism in secret, using three mirrors and a witness from the Quiet Seal. Requires breaking into the prison and holding the demon at bay while the ritual completes. \textit{Cost:} The demon is sealed, but one of the witnesses (the party's choice) forgets a significant memory---the face of a loved one, the name of a ship, the reason they started walking this road.
  \item \textbf{Break the Girl Out:} The party stages a prison break, smuggling Livia to a safe house in the Vilikari marches. The demon follows. The party must now manage a fugitive arc: feeding the demon's hunger with small lies, finding a true exorcist, and staying ahead of both the Inquisition and the Censor's hired blades. \textit{Cost:} Livia will never fully trust the party. She always wonders if they are saving her or using her as bait.
\end{enumerate}

\textbf{Timers:}
\begin{itemize}
  \item \textbf{Inquisition's Schedule [6]} --- ticks each day. When filled, the girl is burned at the Triumph. The demon escapes into the crowd.
  \item \textbf{Demon's Grip [8]} --- ticks each time Livia speaks in a forgotten tongue, each time a witness looks away. When filled, the demon permanently possesses her. She becomes a vessel. Fabius will not survive the guilt.
  \item \textbf{Mirror Integrity [4]} --- ticks each time the prison's silvered mirrors crack. When filled, the demon walks through the reflections. The prison becomes a charnel house.
\end{itemize}

\emph{These cases are not solved. They are \textbf{deferred}. If you succeed where I failed, write to me. I will add your name to the list of those who closed a door I could only hold ajar.}

\begin{flushright}
Saikou Ira\\
Port Shed, Dhaharan Coast\\
The ink is ash. The mirror is still waiting.
\end{flushright}