\appendix
\part{Appendices}

\section{Field Procedures}

\emph{For the use of investigators, exorcists, and those who walk the edge.}

\subsection{The Silver Test}

Carry a small silver coin. When you suspect a supernatural presence, place the coin on your tongue. If the coin tastes of copper or blood, you are in the presence of a vampire, lycanthrope, or demon. If the coin remains tasteless, the threat is spiritual (ghost, poltergeist, Anchor) rather than physical.

\emph{Do not swallow the coin.}

\subsection{The Mirror Array (The Witness)}

To seal a room against possession or haunting:
\begin{enumerate}
  \item Place a silvered mirror facing each doorway.
  \item Place a mirror facing each window.
  \item Place a single mirror in the center of the room, facing the ceiling.
\end{enumerate}
Clean the mirrors with salt water before placement. The central mirror serves as a \emph{focus}---any spirit that attempts to enter will see its own reflection and hesitate.

\subsection{The Traveler's Shortcut}

When you need to bypass a patrol, a locked door, or a stretch of dangerous road, hold a pinch of dust from a crossroads in your left hand and whisper: \emph{``The road that was not there, I walk it now.''} Throw the dust into the air. If The Traveler is listening, a path will open. It will not last long. Do not hesitate.

\subsection{Raéyn's Desperation}

If you are drowning or trapped beneath the waves, speak Raéyn's name three times and touch silver to salt water. She may grant you a single breath---enough to reach the surface or to strike one blow. This favor is never forgotten. She will ask for something in return, usually on a night when the tide is wrong and the wind is silent.

\subsection{The Sentinel's Ward (Exorcism)}

To exorcise a possessing demon:
\begin{enumerate}
  \item Draw a circle of salt and iron filings around the possessed. The circle must have \emph{three} witnesses: one who loves the victim, one who fears the demon, and one who is neutral.
  \item Place three mirrors: one facing the victim, one facing the demon (reflected in the victim's eyes), and one facing the exorcist.
  \item Speak the demon's true name three times, learned through The Witness's revelation.
  \item On the third utterance, the demon is forced into the central mirror.
  \item Shatter the mirror with a silver hammer. The Sentinel seals each shard.
\end{enumerate}
\emph{The demon is not destroyed. It is distributed. It cannot act, but it may be reassembled by a skilled summoner. Dispose of the shards in separate bodies of water.}

\subsection{The Witness's Knot}

When interviewing a suspect who may be possessed or cursed, tie a red thread around your left wrist and the suspect's right wrist. Explain that the knot binds the truth. Speak your questions clearly. The suspect cannot lie while the knot holds---but the knot will tighten if the truth is painful. Cut the knot when the interview concludes. Do not reuse the thread.

\subsection{The Salt Line}

A continuous line of salt will bar most spirits and lesser cursed creatures. Use sea salt for Linn creatures, rock salt for Aeler entities, and salt blessed by any faith for general purposes. A single break---a finger's width---is enough for a determined spirit to cross.

\subsection{The Ninth Measure}

When you encounter a creature or spirit that seems to have no weakness, look for what is \emph{missing}. The ninth cup un-poured. The ninth bell un-rung. The ninth name unspoken. The ninth step uncounted. Fill the absence, and the creature will sometimes cease. \emph{I do not know why this works. I do not ask.}

\clearpage
\section*{The Three Calls: An Exorcism Deduction Mini-Game}
\addcontentsline{toc}{section}{The Three Calls: An Exorcism Deduction Mini-Game}
\noindent\textcolor{witnessSilver}{\titlerule[0.8pt]}
\vspace{0.3em}

\noindent\textbf{Core Concept:} The demon’s true name is its \textit{first lie}—the clause it hid, the violence it used as a key, the moment the host said ``yes'' to something they should have refused. Players must uncover three or more identifying \textit{traits} through research and investigation, then speak the name during a three‑stage exorcism ritual. Each confirmed trait slows the Possession Depth timer; incorrect guesses or failed rolls accelerate it. Success seals the demon; failure gives it control.

\subsection*{Pre‑Ritual Investigation (Untimed)}
Before drawing the circle, players gather clues via research, witness interviews, or forensic examination. Each discovered trait reduces the demon’s ability to lie. \textbf{Three confirmed traits} uniquely identify a demon (no two demons share the same set). Traits can also be confirmed during Step 1 (Draw Circle) by spending an action on research (see Timer Triggers).

\begin{table}[h!]
\centering
\caption{Trait Discovery Mechanics}
\begin{tabular}{@{}l p{10cm}@{}}
\toprule
\rowcolor[gray]{0.9}
\textbf{Action Type} & \textbf{Effect} \\
\midrule
\textbf{Investigation} (during ritual) & Roll Presence+Command vs. DV 12. Success = 1 trait; Critical = 2 traits; Failure = nothing (+1 timer tick if during ritual) \\
\textbf{Research} (pre‑ritual) & No time cost; reveals traits through records, interviews, observation. \\
\textbf{Weakness Test} & Attempt suspected weakness (e.g., holy water, salt circle, iron). Success confirms trait. \\
\bottomrule
\end{tabular}
\label{tab:trait-discovery}
\end{table}

\subsection*{The Exorcism Timer : Possession Depth}
Starts at \textbf{0/8} when the circle is drawn. \textit{Only advances on specific actions—research buys time; guessing costs it.}

\begin{table}[h!]
\centering
\caption{Timer Triggers \& Effects}
\begin{tabular}{@{}l p{9cm}@{}}
\toprule
\rowcolor[gray]{0.9}
\textbf{Trigger} & \textbf{Effect} \\
\midrule
Failed Investigation \textit{during} ritual & +1 segment; demon whispers through host. \\
\textbf{Call attempt with < required traits} (1st call needs 1, 2nd needs 2, 3rd needs 3) & \textbf{+1 segment}; demon mocks the guess (\textit{no roll made}). \\
Call attempt with correct name but insufficient traits & +1 segment; proceed to roll (treat as failed trait check). \\
Call attempt with wrong name ($\geq$ required traits) & +2 segments; host convulses; 1 witness looks away. \\
Call attempt with correct name ($\geq$ required traits) & 0 segments; proceed to roll. \\
1 Action spent researching \textit{during} Step 1 (Draw Circle) & –1 segment; gain 1 trait (if undiscovered). \\
Call roll result: Miss (0 successes) & +1 segment; demon gains 1 thread / breaks eye contact (per call). \\
Call roll result: Partial ($0 < \text{successes} < \text{DV}$) & 0 segments; \textit{no timer advance} (progress without success). \\
\bottomrule
\end{tabular}
\label{tab:timer}
\end{table}

\subsection*{Resolving the Three Calls}
Replace SPIRIT+Lore with \textbf{Presence+Command} rolls. Each call requires a minimum number of confirmed traits to attempt; otherwise the demon mocks the guess and the action ends immediately (timer ticks as above).

\begin{enumerate}[leftmargin=*]
\item \textbf{First Call} – \textit{``I name you [TRUE NAME]. You are seen. You are witnessed.''}
  \begin{itemize}
    \item \textit{Trait check:} $\ge$1 trait?  
      \begin{itemize}
        \item \textbf{No} $\rightarrow$ +1 segment, demon mocks, \textbf{end action} (no roll).
        \item \textbf{Yes} $\rightarrow$ Roll Presence+Command vs. DV (see table).
          \begin{itemize}
            \item \textit{Success:} Proceed to Second Call.
            \item \textit{Fail:} Demon gains 1 thread; \textbf{+1 segment} (from miss).
          \end{itemize}
      \end{itemize}
  \end{itemize}
\item \textbf{Second Call} – \textit{``I name you [TRUE NAME]. You are bound. You are contained.''}
  \begin{itemize}
    \item \textit{Trait check:} $\ge$2 traits?
      \begin{itemize}
        \item \textbf{No} $\rightarrow$ +1 segment, demon mocks, \textbf{end action}.
        \item \textbf{Yes} $\rightarrow$ Roll Presence+Command vs. DV.
          \begin{itemize}
            \item \textit{Success:} Proceed to Third Call.
            \item \textit{Fail:} Host gains 1 FATIGUE; demon breaks eye contact; \textbf{+1 segment}.
          \end{itemize}
      \end{itemize}
  \end{itemize}
\item \textbf{Third Call} – \textit{``I name you [TRUE NAME]. You are \textbf{unmade}.''}
  \begin{itemize}
    \item \textit{Trait check:} $\ge$3 traits (unique identification).
      \begin{itemize}
        \item \textbf{No} $\rightarrow$ +1 segment, demon mocks, \textbf{end action}.
        \item \textbf{Yes} $\rightarrow$ Roll Presence+Command vs. DV.
          \begin{itemize}
            \item \textit{Success:} Demon mirror shatters; shards fly inward.  
              \begin{itemize}
                \item \textit{Normal success:} Demon trapped in exorcist’s mirror.  
                \item \textit{Critical success (all 4 traits confirmed):} Witness whispers the summoner’s name.
              \end{itemize}
            \item \textit{Fail:} Demon possesses exorcist mirror (per core rules); \textbf{+1 segment}.
          \end{itemize}
      \end{itemize}
  \end{itemize}
\end{enumerate}

\subsection*{Failure States}
If the Possession Depth timer reaches \textbf{8/8}:
\begin{itemize}
  \item \textbf{Mirror Integrity Failure First:} If any mirror hits 0 Integrity (per core rules), it shatters – demon controls that reflection.
  \item \textbf{Timer Hits 8/8:} Demon achieves \textit{total breakthrough}:
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=red!60!black,title={Total Breakthrough}]
The host’s eyes glow with the demon’s true form. Silver mirrors fog, then crack inward as if sucked into glass. The demon speaks through the host’s mouth—not in lies, but in the awful truth of its first lie. It gains control of the host’s body and one mirror’s reflection (usually the exorcist’s). The ritual fails; the demon must be re‑bound.
\end{tcolorbox}

\begin{itemize}
  \item \textbf{Consequences:} Host remains possessed (+2 DV on future exorcism rolls); one mirror permanently shattered; each witness who looked away gains 1 Trauma.
\end{itemize}

\subsection*{Demon Trait Master List}
\textit{13 distinct demons. Players need $\ge$3 traits to uniquely identify. Traits are discovered via pre‑ritual research or during Step 1 (Draw Circle).}

\begin{table}[h!]
\centering
\caption{Demons \& Traits (Player‑Facing)}
\begin{tabular}{@{}l l p{5.5cm}@{}}
\toprule
\rowcolor[gray]{0.9}
\textbf{Demon} & \textbf{Class} & \textbf{Traits (Discoverable Evidence)} \\
\midrule
Harperar & A & Invisible contract; Ledger: +1 debt; Offers power via solo bargain; \textit{Weak to Runekeeper enforcement} \\
Kragoth & B & Scorch‑mark door; Bell: rings per corpse; Erases, doesn’t speak; \textit{Requires 3+ unwavering witnesses} \\
Malenka & C & Favoured food = ash; Mother’s face blurs; Wakes with hands on lover’s throat; \textit{Requires faith‑matched witness} \\
Ghoultide & D & Fear grows teeth; Grief walks into next room; Ambition starves the body; \textit{Cured by sensory deprivation} \\
Nyx & A & Contract: “Sign here”; Ledger: debt doubled overnight; Promises knowledge for secrets; \textit{Contract void if 3rd party witnesses} \\
Zaltar & B & Bell: tolls at midnight; Scorch: shape of open grave; Drains life from nearby plants; \textit{Weak to running water} \\
Vorlagh & C & Host’s laughter = hollow; Forgets own name; Sees doppelgänger in mirrors; \textit{Weak to holy water of host’s faith} \\
Mirelle & D & Joy turns to dread; Hope becomes obsession; Trust curdles into paranoia; \textit{Cured by confessing a secret sin} \\
Ravik & A & Contract: blank page; Ledger: debt in blood, not coin; Grants visions of future wealth; \textit{Contract burns if spoken aloud} \\
Skald & B & Bell: one toll = one lie; Scorch: shaped as a lying tongue; Freezes shadows in place; \textit{Weak to loud, sustained noise} \\
Mira & C & Host’s voice echoes; Dreams feel like memories; Finds unfamiliar objects in hand; \textit{Weak to laughter} \\
Thorne & D & Sadness = physical weight; Anger = invisible fire; Greed = hunger that never fills; \textit{Cured by fasting 3 days} \\
Vyre & A & Contract: sealed with tear; Ledger: debt paid in favours; Offers revenge for a whisper; \textit{Weak to moonlight on silver} \\
Grim & B & Bell: tolls for the living; Scorch: shaped as an open eye; Makes flowers bloom then rot; \textit{Weak to buried iron} \\
\bottomrule
\end{tabular}
\label{tab:demons}
\end{table}

\subsection*{Example Playthrough: The Censor’s Daughter (Ecktoria, 859 AR)}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title=The Censor’s Daughter]
\textit{Pre‑Ritual Investigation (Untimed):}  
Players checked the censor’s ledgers (found \textbf{impossible debt}), interviewed villagers (learned he paid for his wife’s medicine through a shady loan – \textbf{invisible contract}), and noted the daughter’s aversion to bread (\textbf{favourite taste like ash}). Cross‑referencing loan records showed the debt was sold to a deceased collector (\textbf{debt that could never be paid}). \textbf{Three traits confirmed for Harperar (Class A): Invisible contract, Ledger debt, Impossible debt.}

\textit{Ritual Execution:}
\begin{itemize}
  \item \textit{Step 1 (Draw Circle):} No research needed (pre‑done); timer starts at 0/8.
  \item \textit{Step 2 (Place Mirrors):} Completed.
  \item \textit{First Call:} “I name you Harperar. You are seen. You are witnessed.” ($\ge$1 trait) → Success. Host convulses; demon mirror shows a ledger‑clutching figure. Timer : 0/8.
  \item \textit{Second Call:} “I name you Harperar. You are bound. You are contained.” ($\ge$2 traits) → Success. Demon mirror sweats; condensation drips like ink. Timer : 0/8.
  \item \textit{Third Call:} “I name you Harperar. You are \textbf{unmade}.” ($\ge$3 traits) → Success. Demon mirror shatters; shards fly inward. Demon trapped in exorcist’s mirror.
\end{itemize}
\textit{Outcome:} The girl is freed. The censor confesses to loan‑sharking. Shards are dispersed to river, well, and ink bottle. The timer never ticked – preparation won the day.
\end{tcolorbox}

\noindent\textcolor{witnessSilver}{\textit{--- The first lie is the key. The timer is the enemy. ---}}