\section*{Saikou Ira – Field Equipment}

\noindent\textit{“I carry no codex. I carry witnesses.”}

\begin{itemize}[leftmargin=*]
  \item \textbf{Silver Mirror (cracked on the left edge)}: Used for the Mirror Array and exorcisms. The crack appeared during an exorcism in Ecktoria. Saikou does not remember which one.
  \item \textbf{Censer (brass, inscribed with the Sigil of the Witness)}: Burns a compound of saltpeter, silver filings, and ironwood resin. The smoke reveals demonic possession and lingering Echoes.
  \item \textbf{Silver Coin (worn smooth)}: For the Silver Test (placed on the tongue). It tastes of copper when a vampire, lycanthrope, or demon is near.
  \item \textbf{Bell (bronze, no clapper)}: Rung to attract the attention of lesser spirits or to mark the boundary of a containment circle. The clapper was lost decades ago; the bell still tolls for Saikou alone.
  \item \textbf{Small Leather Pouch (salt and iron filings)}: Used to draw containment circles and ward thresholds. The salt is from the Mistlands; the iron was filed from a set of manacles that once held a possessed child.
  \item \textbf{Journal (waterproofed with eel‑fat)}: Contains case notes, autopsy sketches, and the true names of three Class C demons. The pages are numbered in eights; the ninth page of each entry is left blank.
  \item \textbf{Thorn‑Pin (silver)}: Used to prick a possessed subject’s finger. A drop of blood on the pin reveals the demon’s true name as a faint whisper. The pin was a gift from a Daughter of the Cord – a debt he has not yet paid.
  \item \textbf{Knife (curved, no reflection)}: For cutting cords, opening sealed letters, and, when necessary, cutting short an exorcism gone wrong. The blade does not show in mirrors; Saikou never uses it on the living.
  \item \textbf{Moonsteel Cutlass (curved, no reflection)}: A sword that culls. A sword that \textit{whispers.}
\end{itemize}

\gray{I once heard of a Trow who faced three mirrors at a crossroads at dusk. He was forced to trade his own name. Now he wanders, nameless.}

\clearpage
\part{The Thirteenth Chime}

\section{Conclusion: On the Necessity of Witness}

I have written this compendium not to frighten, but to \emph{prepare}. The creatures I have described are not myths. They are neighbors, patients, clients, and---sometimes---the faces we see in the mirror when we are alone at midnight.

The Witness has taught me that truth is not a weapon. It is a \emph{tool}. The truth reveals the weak point, the bargain, the memory that binds the spirit to its hunger.

I have killed. I have saved. I have wept over bodies I could not help and celebrated over curses I could not break.

The work continues.

If you are reading this, you are now a witness. The creatures I have described are real. They are hungry. They are waiting.

But so are we.

The ninth cup is not poured. The ninth bell is not rung. The threshold is held---by you, by me, by the quiet ones who tend the lamps and count the breaths and remember the names that others have forgotten.

Do not look away.

\begin{flushright}
Saikou Ira\\
Port Shed, Dhaharan Coast\\
The mirror is clean. The truth is waiting.
\end{flushright}

