\part{The Cursed}

\section{Lycanthropes}

The werewolf is a creature of paradox: a curse that feels like freedom, a transformation that erases the self, a hunger that is also a family. I have tracked lycanthropes across three continents. I have killed two. I have saved four.

\subsection{Origins and Cultural Variations}

Lycanthropy is a \emph{spiritual infection}---a fragment of Malachai's chained essence passed from host to host through violence, but only if the victim is in a state of profound emotional vulnerability.

\begin{description}
  \item[Linn Sea-Wolf:] Transforms into a massive seal or walrus. Solitary, melancholic, preys on ships.
  \item[Ykrul Sky-Hound:] Hawk-headed hunter that transforms during the day. Hunts oath-breakers, not flesh.
  \item[Valewood Root-Walker:] Transforms into a tree—rooted, immobile, but all-seeing. Guardians, not predators.
  \item[Aeler Stone-Hide:] Massive, slow, nearly invulnerable. Transforms only underground. A mining hazard.
\end{description}

\subsection{Investigation Notes}

\begin{itemize}
  \item Silver through the heart or brain kills them.
  \item Aconite (wolfsbane) poultice applied within the hour can prevent transformation.
  \item The full moon is not a universal trigger; transformation responds to \emph{emotional} cycles.
  \item The Well of Ichor can purge lycanthropy, but the journey to Zahk'Kozahn is itself a trial.
\end{itemize}

\textbf{The Witness's Observation:} Lycanthropes who remember their human lives clearly are less dangerous than those who do not. The memory of love acts as a leash on the beast.

\textbf{Case Study: \emph{The Hunters of the Scarlet Wood}} (Crimson Valley, 862). A pack of feral lycanthropes with no memory. I placed a mirror at the cave entrance. As the first wolf emerged, it saw its own reflection---and for a moment, I saw recognition. That wolf was a young silk-weaver who had gone missing two years prior. I spoke her name. She stopped. We negotiated a quarantine. She now feeds on deer and speaks to her daughter through a silvered glass.

\section{Vampires}

Vampires are the eldest children of Malachai, the first to drink from the Chained Angel's wound.

\subsection{The Thirteen First Vampires}

I have met three of the Thirteen. I will describe them briefly, as a warning.

\begin{description}
  \item[Baron Quentin Asselin:] The first to drink. Disillusioned, melancholic, almost gentle. Converts guests or drives them mad. \emph{Never accept his wine.}
  \item[Sister Maris:] Runs a children's hospice. The children never leave. The Everflame has tried to destroy her hospice four times. Each time, the children defended her.
  \item[The Sparrow:] Kills only the wicked, or so he claims. Cannot look at mirrors. Has saved my life twice. I suspect he is merely \emph{bored}.
\end{description}

\subsection{Lesser Bloodlines}

\begin{itemize}
  \item \textbf{Linn Draugr:} Breathe underwater, trapped by sea-salt circles.
  \item \textbf{Aeler Stone-Sleepers:} Drink breath, awakened by bronze bells at dawn.
  \item \textbf{Ykrul Bone-Eaters:} Turn to dust when killed, but rise if their name is spoken. Erase the name.
  \item \textbf{Vhasian Candle-Wights:} Extinguish flames through windows. A lit candle at the threshold bars them.
  \item \textbf{Lethai-ar Silk-Drinkers:} Feed on desire. Vulnerable to vows witnessed by a third party.
\end{itemize}

\textbf{The Sentinel's Ward:} When I must enter a vampire's lair, I invoke The Sentinel to bless the threshold of my mind. A silver mirror worn over the heart, inscribed with the sigil of the Sealed Gate, prevents the vampire's hypnotic gaze from finding purchase. I have walked into the dens of three vampires wearing that mirror. Only once did I have to run.

\textbf{Case Study: \emph{The Court Painter of Mbaro}} (Oshiira, 856). A Silk-Drinker feeding on nobles' ambition. I confronted him in the Widow's weigh-house. He did not flee. He wept and confessed. The Widow commuted his death sentence to indefinite confinement in a silver-lined cell. He paints the Widow's portrait each year. She has not aged in seven years. I do not know if this is related.

