¥part{The Hungry}

¥section{Ghouls and the Flesh竏脱aters}

The ghoul is not a single creature, but a ¥emph{state}竏停娣竏誕 progression from unquiet death to active predation. I have encountered ghouls in every region of the Amaranthine, from the frozen barrows of Linn to the sun竏鍛aked crypts of Ashaan. Their origins are varied, but their behavior is remarkably consistent.

¥subsection{The Three Origins}

¥textbf{Origin One: The Starved Anchor}

An Anchor that is denied resolution for too long may begin to consume the life force of the living to sustain its own existence. These ghouls retain their intelligence and memory. They are the most dangerous, because they are the most ¥emph{pitiable}.

¥textbf{Origin Two: The Corpse竏巽ed}

A body buried in ground that is rich in death竏停娣竏誕 battlefield, a plague pit, a site of repeated sacrifice竏停娣竏知ay rise as a ghoul regardless of the soul's disposition. These ghouls are ¥emph{reanimated corpses}, not spirits. They have no memory of their former lives. Fire stops them. Decapitation stops them. Nothing else does.

¥textbf{Origin Three: The Cursed}

Malachai's touch takes many forms. The ghoul is one of her simpler curses竏停娣竏誕 gift to those who begged for immortality and received a half竏値ife of hunger instead. These ghouls retain their intelligence, but their morality erodes over time.

¥subsection{Investigation Notes}

¥begin{itemize}
  邃ｹtem Ghouls cannot digest food. They consume living flesh to absorb ¥emph{life essence}.
  邃ｹtem They are vulnerable to silver, but not to the same degree as vampires or lycanthropes. A silver blade wounds them; fire destroys them.
  邃ｹtem A ghoul that has not fed for a month will enter a state of torpor, appearing dead. It can be buried safely, but if disturbed, it will wake ravenous.
  邃ｹtem There is no known cure for ghoulism. The Well of Ichor in Zahk'Kozahn offers release, but the price is memory.
¥end{itemize}

¥textbf{Varnek Karn's Mercy:} When I must destroy a ghoul that was once a good person, I invoke Varnek Karn. A cup of salt water竏停娣竏鍛lessed by a witness who did not partake in the killing竏停娣竏恥oured over the remains after destruction ensures that the soul will not rise again. The water must be thrown by someone who never drew steel against the creature: a bystander, a family member, a priest. Varnek Karn demands that the terms of the bargain be held separate竏停娣竏稚he hand that strikes and the hand that absolves. I do not know why, but I obey.

¥textbf{Case Study: ¥emph{The Dock竏坦at of Port Shed}} (Dhaharan Coast, 860). A young Sidhi woman was infected with ghoulism after drinking from a cursed well. She hid her condition for two years, feeding on stray animals. When she could no longer control the hunger, she fled into the sewers. I found her竏停娣竏痴tarved, weeping, clutching letters to her mother. I negotiated a feeding schedule: animal blood, provided by a local butcher in exchange for her labor as a night watchman (ghouls do not sleep, and their senses are sharp). She has not harmed a human in five years. Her mother knows. The butcher knows. I have told no one else.

