\chapter{Other Entities of the World}

\section*{A Note on What This Chapter Contains}

The preceding chapters taxonomise the dead, the cursed, and the infernal. But there exist entities that do not fit those tidy categories. They are not ghosts, though they may wear the faces of the dead. They are not demons, though they may bargain. They are \textbf{obligations given form}——debts that have learned to walk, promises that have grown teeth. This chapter records four such entities. I have encountered three of them. I have outlived two.

\emph{—— Saikou Ira, Port Shed, the hour before dawn}

\section{The Oath-Keeper(Possessor Variant)}

The Oath-KeeperI described in the Appendix is a spirit of unpaid obligation, formless and inexorable. But there is a rarer, more terrible variant——one that does not simply demand payment. It \textbf{wears the face of the person you failed}.

\subsection{Manifestation}

A Oath-Keeperof this type appears when two conditions are met:

\begin{enumerate}
    \item You owe a significant debt to someone——coin, oath, loyalty, care.
    \item You have knowingly broken that oath or refused payment.
\end{enumerate}

The Walker does not emerge from the Hollow or the Ways Between. It emerges from \textbf{your own memory of the wronged person}. At first, it is only a flicker——their face in a crowd, their voice on the wind. But each time you ignore the debt, the manifestation grows more solid. Eventually, it can walk beside you, speak with their voice, and——if the debt is left unpaid long enough——\textbf{possess the living body of someone who resembles them} or \textbf{assume a physical form indistinguishable from the original} (except under silver mirror or true name).

\subsection{Investigation Notes}

\begin{itemize}
    \item \textbf{The Silver Test} (see Appendix): A silver coin placed on the tongue tastes of \textit{cold iron and spoiled honey}——not copper (vampire) nor blood (demon), but the specific taste of a promise left to rot.
    \item \textbf{The Mirror Test}: The Walker’s reflection is absent——or, worse, shows the \textit{character’s own face} instead of the Walker’s.
    \item The Walker cannot be harmed by any mundane weapon. Silver and fire cause it to recoil for one exchange, but do not banish it.
\end{itemize}

\subsection{Behavior}

The Walker does not attack. It \textbf{accuses}. It speaks in the wronged person’s voice, asking a single question, repeated each time it appears:

\begin{quote}
\emph{“Why did you not keep your word?”}
\end{quote}

It does not threaten. It does not bargain (initially). It simply \textbf{witnesses} your shame. Each time you encounter it and fail to answer (or answer dishonestly), the GM ticks a new timer: \textbf{Debt–Walker’s Patience [6]}.

\subsection{Possession of a Living Host}

If the Debt–Walker’s Patience timer fills, or if the character deliberately seeks out the wronged person to harm them, the Walker may \textbf{possess a living person}——usually someone who reminds it of the original (a relative, a friend, a stranger with the same name). The host’s eyes become flat; their voice takes on the Walker’s cadence. They will not remember the possession afterward.

\textbf{Possession Mechanics:} Treat as a Class C Possessor (see Chapter 4, \textit{Demons and the Infernal}). The Walker can be exorcised using The Sentinel’s Exorcism (Chapter 4, \S 0.5.3), but the exorcist must know the \textbf{name of the wronged person}, not the Walker’s name. The Walker has no name of its own——only the debt it represents.

If the host is killed during possession, the Walker does not die. It simply returns to its formless state and may seek a new host within 1d4 days.

\subsection{Resolution}

The Oath-Keepercannot be banished by force——only by \textbf{payment of the original debt}. Payment must be:

\begin{itemize}
    \item \textit{Coin or kind}, if the debt was material.
    \item \textit{A public apology}, witnessed by a third party, if the debt was an oath.
    \item \textit{A substitute offering}——something of equal emotional weight to what was owed (e.g., a cherished memory, a year of service, the character’s own true name spoken in front of a witness).
\end{itemize}

If the original wronged person is dead, the Walker will accept a \textbf{rite of remembrance} (Spirit + Lore, DV 4) performed at their grave, provided the character speaks aloud the name of the dead and confesses the broken promise.

Once the debt is paid, the Walker vanishes. It may return if the character incurs a new debt with the same person (or another). It bears no grudge——it is not a demon. It is a \textbf{reminder}.

\subsection{Clock: Debt–Walker’s Patience [6]}

\begin{itemize}
    \item \textbf{Tick} when the character encounters the Walker and refuses to answer, lies to it, or harms its host.
    \item \textbf{Tick} when the character incurs a new debt with the same wronged person (or their family).
    \item \textbf{Tick} when the GM spends 2+ Story Beats on the character’s ignored obligations.
\end{itemize}

When the timer fills, the Walker possesses the character \textbf{themselves}——not as a full takeover, but as a \textbf{compulsion} to seek out the wronged person and pay the debt, by any means necessary. The character must succeed on a Spirit + Resolve test (DV 5) to resist acting on this compulsion for one scene. Failure means they will abandon their current objective to pursue payment.

\subsection{Case Study – The Merchant of the Broken Ledger}

\emph{Case No. 862–PS–11. Status: Closed (paid in full).}

A Sidhi factor named Tema had borrowed a considerable sum from a ship captain to fund a spice voyage. The voyage failed. Tema did not repay. The captain died of a fever, and Tema told himself the debt died with him.

Three months later, the captain sat at Tema’s desk.

Tema did not flee. He wept. The captain asked the question once, then again, then a third time. Each time, Tema’s Patience timer ticked. On the fourth tick, Tema sold his ledger——the one that recorded every bribe and betrayal of his career——and paid the coin to the captain’s widow. The Walker nodded and vanished. Tema now works as a scribe for the widow. He says the work is honest. I believe him.

\section{The Counting Wraith (Class I variant)}

\emph{First described in the margins of this compendium (Chapter 1, \S 0.1). Here given formal entry.}

The Counting Wraith appears when someone miscounts nine steps, nine cups, nine breaths. It is not a ghost of a person; it is a ghost of an \textbf{error}. The error is its anchor; the witness is its prey.

\subsection{Manifestation}

A faint chill, a missed heartbeat, a stair that feels one step longer than it should. The Wraith is invisible to the naked eye, but a silver mirror shows its outline——a tall, gaunt figure with long fingers that trail threads of gray light.

\subsection{Investigation Notes}

\begin{itemize}
    \item The Wraith cannot be fought. It can only be \textbf{counted}.
    \item It speaks only once: \emph{“You owe a number.”}
    \item If the victim recites aloud the exact sequence of eight steps they took to cross the threshold where the Wraith appeared, the Wraith recedes. If they miscount, it returns the next night and takes a minor memory (the name of a street, the face of a porter, the taste of bread).
\end{itemize}

\subsection{Resolution}

A correct count. If the count is wrong again, the Wraith takes a \textbf{significant} memory: a name, a face, a year. I have seen a merchant forget his daughter’s name. He did not weep. He could not remember why he should.

\subsection{Clock: Wraith’s Patience [6]}

Ticks each time the victim miscounts. When full, the Wraith takes the victim’s \textbf{true name}. The victim becomes a stranger to everyone who knew them.

\subsection{Prevention}

Always count to eight. Then start again. The ninth breath is not for you.

\section{The Weeping Shade (Class II variant)}

A ghost bound by a secret promise of silence. It cannot speak its unfinished business; it can only weep. The weeping is not sadness——it is the pressure of an unspoken name.

\subsection{Manifestation}

A pale, translucent figure with a hand clasped over its mouth. Tears run down its face but evaporate before they touch the ground. All within Near range suffer –1 die to Empathy and Insight rolls (the weeping is distracting). On a successful Manifestation (DV 3), the Shade touches a target; the target cannot speak for one exchange.

\subsection{Resolution}

The Shade cannot be reasoned with directly; it cannot speak. It can only be \textbf{named}. Someone who knows the secret the Shade is bound to protect must speak it aloud in the presence of a witness. The Shade will hear, nod, and pass.

I have seen a woman confess to a murder that was not her own. The Shade had been bound by the real murderer. The Shade passed; the woman became haunted by a different silence.

\subsection{Clock: Silence’s Weight [6]}

Ticks each time the secret is kept. When full, the Shade manifests permanently in the room where the promise was made. The room becomes a threshold; the threshold becomes a prison.

\section{The Unwitnessed Door (Threshold Entity)}

A door that appears only to those who take the ninth step without meaning to. It leads to a room between places, where two mirrors offer two selves. The room does not guarantee which mirror is which. The Witness and Aveh are not in the habit of labeling their wares.

\subsection{Manifestation}

A door of gray wood, unremarkable except that it stands where no door stood before. Its handle is warm. It cannot be forced open——it opens only when the threshold is \textbf{acknowledged}.

\subsection{Investigation Notes}

\begin{itemize}
    \item Entering the door requires a Spirit + Resolve test (DV 5). On a success, the character reaches the room of mirrors. On a failure, they are pulled into the Hollow for 1d4 hours and emerge with no memory of the experience——but their shadow is ever so slightly taller.
    \item In the room, two mirrors face each other. One is of the Witness (shows the self you are meant to be). One is of Aveh (shows the self you should not be). The mirrors are not labeled.
    \item The character must choose a mirror and speak a name——either their own true name, or the name they believe they are meant to speak. The GM determines the outcome in secret.
\end{itemize}

\subsection{Resolution}

There is no combat. The door closes when the choice is made. The character may never find the door again——or it may appear at every ninth step. I have seen a woman enter the Unwitnessed Door and emerge as a completely different person. I have seen a man enter and emerge as himself, but kinder. I do not know which is which.

\subsection{Clock: The Door’s Memory [6]}

Ticks each time the character uses their new self (if transformed). When full, the door appears again, offering a second choice——or a return to the old self. The cost is a debt: the room will demand something in return for another chance.

\section{Closing: On Witnessing the Unwitnessed}

These entities are not common. You may never encounter a Counting Wraith or an Unwitnessed Door. But if you do, remember: they are not punishments. They are \textbf{interest}. The debt was already there. They merely came to collect.

\emph{—— Saikou Ira, Port Shed, under a borrowed lamp}

\section{The Deep Drake – The Pressure that Thinks}

\emph{Case File No. 871–AE–00. Status: Unresolved. Threat Level: Mythic (Tier V–VI).}

The Aeler do not speak of this creature. They call it \textbf{the Pressure that Thinks}——a name, not a title. To name it otherwise is to invite its attention. It dwells in the deepest galleries beneath Khaz‑Vurim, where the stone sweats black and the air has not been breathed by any living lung for ten thousand years. The Deep Accountants have recorded ninety‑seven expeditions sent to explore those depths. Zero have returned.

I did not learn of the Pressure from the Aeler. I learned from Nidhoggr.

\subsection{The Bargain with Nidhoggr}

I sat three days in the Whispering Vaults with my ear pressed to a vein of obsidian. Nidhoggr does not speak; he \textbf{dreams}. The information came as a vision: a serpent of polished jet, its hide etched with runes that predate the Aeler script, coiled around a heart of living rock. The serpent's eyes were not eyes——they were \textbf{pressure gauges}, measuring the weight of centuries.

In exchange for this vision, Nidhoggr demanded a memory of my own——not a painful one, but a trivial one: the name of my first ship. I gave it. I do not miss it. The ship was called the \textit{Copper Bell}. I had forgotten until I wrote this sentence.

The vision showed me the Pressure's lair, its hunting patterns, and the fate of those it consumes. It also showed me its lineage. The Pressure is not Nidhoggr. It is his \textbf{child}——a wyrm hatched from a scale that fell into the deep during the God‑War, fed on the psychic residue of millennia of Aeler breath‑counting. Where Nidhoggr dreams of the past, the Pressure \textbf{dreams of hunger}.

\subsection{Manifestation and Lure}

The Pressure does not attack. It \textbf{invites}.

Miners in the deep galleries sometimes hear a voice——not in their ears, but in their breath. \emph{“Count to nine.”} If they refuse, the voice persists. If they count the ninth breath, the air around them grows heavy. The stone exhales. They begin to walk downward, not by choice, as if the mountain itself has tilted.

These are the Lured. They walk for hours, days, shedding equipment, shedding names, until they reach the Pressure’s lair. There, the Pressure \textbf{consumes} them——not their bodies, but their \textbf{breaths}. Each victim adds a century of accumulated breath to the Pressure’s hoard. Their bodies are left intact, hollow, with eyes that reflect the dark.

\textit{I have seen one such body. Its mouth was open, frozen in the middle of counting. It had reached eight.}

\subsection{Investigation Notes – Reported Capabilities}

All information below is second‑hand, drawn from Nidhoggr’s visions and the testimony of a single survivor (a Vent‑Prior who lost her left arm and her name——she now calls herself \textit{Eight}). Use with caution.

\begin{itemize}
    \item \textbf{Psionic Pressure Aura (Passive):} Any creature within the Pressure’s lair suffers –2 dice to all rolls involving breath, counting, or rhythmic actions. The air itself feels heavy, as if the mountain is leaning on your chest.
    \item \textbf{Phase One – The Counting (Lair as Trap):} The Pressure does not engage directly. Instead, it forces intruders to navigate its lair while \textbf{counting their breaths}. Each character must track a \textbf{Breath Timer [8]}. Ticks advance each time they speak, exert themselves, or fail a Resolve test. When a character’s Breath Timer fills, they become Lured and must follow the Pressure’s voice toward the central chamber, unable to resist unless an ally physically restrains them (requires a Body + Athletics test, DV 4).
    \item \textbf{Phase Two – The Hungry Thought (Psychic Assault):} When at least two characters are Lured, the Pressure manifests as a swirling column of dark, particulate stone. It attacks with \textbf{psionic pressure waves}: targets in Near must test Spirit + Resolve (DV 5) or take Harm 2 (Psychic) and gain the Condition \textit{Echo of the Deep} ( –1 die to all actions until they receive fresh air or a bell is rung).
    \item \textbf{Phase Three – The Uncoiling (Physical Manifestation):} When reduced to 4+ Harm, the Pressure sheds its stone husk and reveals its true form: a serpent of polished jet, thirty feet long, with a crown of brass spikes that hum with psychic resonance. It gains the following abilities:
        \begin{itemize}
            \item \textit{Brass Crown Pulse (Recharge 4–6)}: All characters in Near must test Spirit + Resolve (DV 6) or lose 1 Fatigue and forget a random skill (temporary, lasts until they rest). The Pressure feeds on the memory of the skill.
            \item \textit{Coil of Oblivion}: A single target within Close is grappled (opposed Body + Athletics vs. Pressure’s Body 6). While grappled, the target cannot breathe and takes 1 Harm per round (suffocation). The Pressure will release if it takes 2+ Harm in a single exchange.
        \end{itemize}
    \item \textbf{Phase Four – The Ninth Breath (Desperation):} At 2 Harm or less, the Pressure shrieks. Every character must make a final Resolve test (DV 6). On a failure, they are forced to count the ninth breath aloud. Doing so seals their fate: they become a permanent part of the Pressure’s hoard, unable to leave the lair until the Pressure is destroyed. A character can resist by spending 2 Boons and taking 1 Harm (Psychic) – they bite through their own cheek, the pain grounding them.
\end{itemize}

\subsection{Lair – The Heart of the Ninth Gallery}

The Pressure’s lair is a vast, domed chamber of black basalt, lit only by the faint bioluminescence of fungi that grow on the bones of the Lured. The walls are carved with nine tally marks – repeated thousands of times, each set of nine slightly different. The floor is uneven, scattered with the equipment of lost miners. The air is cold, still, and heavy with the scent of iron.

\subsubsection{Lair Actions (Every Round, Choose One)}
\begin{itemize}
    \item \textbf{Stone Memory}: A section of the floor reverts to its ancient state – a pit or a spike of rock. Characters in that zone must test Body + Athletics (DV 4) or take Harm 1.
    \item \textbf{Pressure Wave}: The chamber’s air pressure drops suddenly. All characters lose 1 Fatigue (the air is sucked from their lungs). Those already at Fatigue 3+ take 1 Harm instead.
    \item \textbf{Witness of the Lured}: A spectral figure appears – one of the consumed miners. It points at a character and whispers the number they last counted aloud. The character must test Resolve (DV 4) or be Shaken for one exchange.
\end{itemize}

\subsection{Counter – What the Aeler Learned (Too Late)}

The Aeler have only one recorded success against the Pressure. A Vault‑Queen ordered a bell of solid copper, forged in silence, and placed it at the entrance to the Ninth Gallery. When the bell was rung, the note was not a note——it was a \textbf{cancelation}. The Pressure retreated for a generation. The bell was later stolen by an Edgewalker who wanted to sell it. The Edgewalker was never found.

\begin{itemize}
    \item \textbf{Copper Bell}: A bell forged in absolute silence, without counting the strikes of the hammer. When rung within the Pressure’s lair, all characters gain +1 die to Resolve tests, and the Pressure’s Brass Crown Pulse cannot be used for the next round.
    \item \textbf{True Name}: The Pressure’s true name is the number that comes after nine. No one knows it. Nidhoggr refused to tell me. He said, \emph{“That is not a name. It is a wound.”}
    \item \textbf{Sacrifice of Breath}: A character may voluntarily hold their breath for three exchanges (Body + Endurance, DV 5 each exchange). If they succeed, they may offer those three breaths to the Pressure as a tithe. The Pressure will cease all attacks for one round and listen. During that round, any character may make a single unopposed action (e.g., ring the copper bell, seal a ward, or run for the exit).
\end{itemize}

\subsection{Aftermath – The Debt to Nidhoggr}

I gave Nidhoggr the name of my first ship in exchange for this knowledge. The ship is gone. The debt is not. Nidhoggr did not ask for payment; he \textbf{recorded} the transaction. One day, he will ask for something else. The Pressure’s child may not be the only worm in the deep.

\emph{If you seek the Pressure that Thinks, bring a copper bell, a companion who can count to eight without flinching, and a will to lose something you did not know you had.}

\emph{—— Saikou Ira, Port Shed, after a dream of obsidian and brass}
