\chapter{Resource Management}

\section{Supply Clock}\index{Resources!Supply Clock}

The Supply Clock is a shared condition for the entire party, representing food, water, and basic gear.\index{Supply}

\subsection{States}\index{Supply!States}

\begin{description}
\item[Full Supply (0 filled)] The party is well-equipped.\index{Supply!Full}
\item[Low Supply (2 filled)] Minor narrative complications: bland food, damaged arrows, thinning waterskins.\index{Supply!Low}
\item[Dangerously Low (3 filled)] Each character gains Fatigue.\index{Supply!Dangerously Low}
\item[Out of Supply (4 filled)] Severe penalties; starvation, dehydration, failing gear.\index{Supply!Out of Supply}
\end{description}

\subsection{Filling the Clock}\index{Supply!Filling}

\begin{itemize}
\item Harsh travel or lost pack animals (GM fiat)\index{Supply!GM Fiat}
\item GM spends 2+ Story Beats\index{Supply!Story Beats}
\item The party chooses to travel light for advantage\index{Supply!Travel Light}
\end{itemize}

\subsection{Emptying the Clock}\index{Supply!Emptying}

\begin{itemize}
\item Reaching civilization resets to Full\index{Supply!Civilization}
\item Foraging/hunting: group Survival check clears 1 segment\index{Supply!Foraging}
\item Downtime in safety removes 1 segment\index{Supply!Downtime}
\end{itemize}

\section{Fatigue}\index{Resources!Fatigue}

Fatigue represents exhaustion, hunger, and strain.\index{Fatigue}

\subsection{Effects}\index{Fatigue!Effects}

\begin{description}
\item[1 Fatigue] Minor drain. On your next roll, re-roll one success (player's choice).\index{Fatigue!1 Level}
\item[2 Fatigue] Worn down. On each roll, re-roll one success.\index{Fatigue!2 Levels}
\item[3 Fatigue] Failing fast. On each roll, re-roll two successes.\index{Fatigue!3 Levels}
\item[4 Fatigue] Collapse/KO/Spiritual break. You fall out of the scene until treated or rescued.\index{Fatigue!4 Levels}
\end{description}

\subsection{Clearing Fatigue}\index{Fatigue!Clearing}

A night's rest with adequate Supply removes 1 level. You cannot clear Fatigue if the party's Supply clock is Dangerously Low or Empty.\index{Rest}

\section{Gear Condition}\index{Resources!Gear Condition}

Gear does not have hit points. It suffers only when drama demands it.\index{Gear}

\subsection{Compromised Items}\index{Gear!Compromised}

\begin{itemize}
\item Introduced via Story Beats or narrative consequence\index{Gear!Complications}
\item A Compromised item gives --1 die on relevant rolls\index{Gear!Penalty}
\end{itemize}

\subsection{Breaking Point}\index{Gear!Breaking}
If a Compromised item suffers another setback, it breaks entirely.\index{Gear!Breakage}

\subsection{Repair}\index{Gear!Repair}

\begin{description}
\item[Field Repair] Temporary; Craft or Survival check removes penalty for one scene\index{Gear!Field Repair}
\item[Proper Repair] Permanent; requires tools, materials, and downtime\index{Gear!Proper Repair}
\end{description}

\section{Asset and Follower Management}\index{Resources!Assets and Followers}

\subsection{Followers (On-Screen)}\index{Followers}

\begin{description}
\item[Cost] Cap$^2$ XP\index{Followers!Cost}
\item[Assist Dice] When applicable, the follower adds help dice equal to \textbf{min(Cap, the helper's relevant Skill)}, capped at +3 dice. Exception: The "Exceptional Coordination" Talent allows one follower to provide +4 assist dice.\index{Followers!Assist Dice}
\item[Cap] 1-5 (5 is exceptional)\index{Followers!Capability}
\item[Upkeep] Each Downtime, pay XP equal to Cap or spend a Scene tending the relationship.\index{Followers!Upkeep}
\item[Risk] If the GM spends 2+ Story Beats on an action you take with assistance, they may endanger, injure, or separate the follower instead of you if fictionally appropriate.\index{Followers!Risk}
\item[Off-Screen Capability] Once per downtime, a Cap 5 follower can solve one significant problem but generates 1 SB for party. The GM must describe how the follower's action creates story consequences for the SB generated.\index{Followers!Off-Screen}
\end{description}

\subsection{Follower Assist}\index{Followers!Assist}
\begin{itemize}
    \item Assist dice come from the helper, not the leader.\index{Mechanics!Assist Dice}
    \item Total Assist on any roll (from any sources) remains hard-capped at +3. Exception: The "Exceptional Coordination" Talent allows one follower to provide +4 assist dice.\index{Mechanics!Assist Maximum}
    \item Slot Limit: Only one follower may assist a given action.\index{Followers!Slot Limit}
\end{itemize}

\subsection{Loyalty \& Bonds (Optional)}\index{Followers!Loyalty}
\begin{itemize}
    \item Track a simple Loyalty tag per follower: Wary / Steady / Devoted.\index{Followers!Loyalty Tags}
    \item Devoted followers can once per arc convert one GM Complication targeting them into a lesser setback.\index{Followers!Devoted}
    \item Wary followers cost +1 XP to Maintain.\index{Followers!Wary}
\end{itemize}

\subsection{Stress, Harm, \& Loss (GM Tools)}\index{Followers!Stress}
\begin{itemize}
    \item \textbf{Pin}: The follower is separated/boxed out.\index{Followers!Pin}
    \item \textbf{Wound}: The follower is Injured: until treated off-screen, their Cap counts as 1 lower.\index{Followers!Wound}
    \item \textbf{Burn}: Mark Neglected immediately.\index{Followers!Burn}
    \item \textbf{Seize}: Escalate to Compromised.\index{Followers!Seize}
    \item \textbf{PC Choice Lever}: The GM should offer the player a save.\index{Followers!Player Choice}
\end{itemize}

\subsection{Off-Screen Assets}\index{Assets!Off-Screen}

\begin{description}
\item[Minor (4 XP)] Safehouse, small charter\index{Assets!Minor}
\item[Standard (8 XP)] Noble title, guild section, spy ring\index{Assets!Standard}
\item[Major (12 XP)] City license, regional network\index{Assets!Major}
\end{description}

\subsection{Asset Activations}\index{Assets!Activations}
\begin{itemize}
    \item \textbf{Off-Screen Activation}: A player may activate an off-screen asset at the very start of a campaign or during Downtime. It costs 1 Boon or 2 XP to activate.\index{Assets!Activation Cost}
    \item \textbf{Off-Screen Effects}: Use each Asset's listed Off-Screen effect once per session for free.\index{Assets!Off-Screen Effects}
    \item \textbf{On-Screen Activations}: To reshape the current scene, spend 1 Boon.\index{Assets!On-Screen Effects}
    \item \textbf{Plausibility Test}: The Asset must have scope and reach.\index{Assets!Plausibility}
\end{itemize}

\subsection{Condition Tracks}\index{Assets!Condition Tracks}

\begin{description}
\item[Maintained] Full capability; no penalties\index{Assets!Maintained}
\item[Neglected] --1 die when used (assist or leverage). Narratively: slower, sullen, short-staffed\index{Assets!Neglected}
\item[Compromised] Unavailable. Narratively: captured, burned, seized, defected\index{Assets!Compromised}
\end{description}

\subsection{Maintenance and Repair}\index{Assets!Maintenance}

\begin{itemize}
\item After a story arc or 2--3 sessions of heavy use, the GM may flag the resource as at risk\index{Assets!Risk Flag}
\item To keep it Maintained, the player must either:
  \begin{itemize}
  \item Spend downtime repairing/servicing it, or\index{Assets!Downtime Repair}
  \item Pay XP immediately (representing replacement parts, masterwork upkeep)\index{Assets!XP Repair}
  \item If neglected, the gear becomes Compromised (--1 die penalty) and may eventually break\index{Assets!Breakage}
  \item Superior and Artifact gear do not require upkeep, but if Compromised through Complications, only narrative quests can repair them\index{Assets!Superior Gear}
  \end{itemize}
\end{itemize}

\subsection{Bond-Driven Resource Generation}\index{Resources!Bond-Driven Generation}
When a player takes a significant action to aid an ally with whom they share a bond, and explicitly references that bond in an intricate description, they may mark that bond to gain 1 boon after the action resolves.\index{Bonds!Resource Generation}

\textbf{Requirements:}
\begin{itemize}
    \item \textbf{Mutual Bond:} Player shares a bond with the ally they're aiding\index{Bonds!Mutual Requirement}
    \item \textbf{Intricate Description:} Player describes how their bond motivates their action\index{Description Ladder!Intricate Action}
    \item \textbf{Significant Aid:} Meaningful assistance (not just +1 die)\index{Followers!Significant Aid}
    \item \textbf{Fiction First:} The bond must genuinely drive the choice to help\index{Fiction First}
\end{itemize}

\textbf{Examples:}
\begin{itemize}
    \item "Remembering how they saved me before, I throw myself in front of the attack meant for them!"
    \item "Thinking of our shared past, I use my last resource to shield them from harm."
    \item "Drawing on our mutual experiences, I rally others to keep fighting alongside them."
\end{itemize}

\textbf{Limitations:}
\begin{itemize}
    \item Once per bond per session\index{Bonds!Session Limit}
    \item Must be a meaningful sacrifice or risk\index{Risk}
    \item GM approval for "significant action"\index{GM Discretion}
    \item Cannot be used for basic assistance rolls\index{Description Ladder!Basic Action}
\end{itemize}

\section{Over-Stack}\index{Combat!Over-Stack}
\begin{itemize}
    \item Structural advantages: active buff/tag, favorable venue/pennant, Follower Initiative unused, on-screen Asset activation, opponent disadvantaged by fiction, ritual prep that applies now.\index{Combat!Structural Advantages}
    \item Trigger: If the party enters a scene with ≥ 3 structural advantages, apply Over-Stack once for that scene: either start one named rail at +1 or the GM banks +1 SB for the first Deck Twist.\index{Combat!Over-Stack Trigger}
\end{itemize}
\end{chapter}
