\chapter{Deck-Based Generators}

\section{Standard Deck Structure}

Fate's Edge uses distinct card-based tools to generate narrative content, each with its own specialized suit meanings and purpose. These generators transform randomness into meaningful story elements while maintaining thematic coherence.

\paragraph{Travel Decks (regional, 52-card).}
\emph{Spade}=Place, \emph{Heart}=Actor, \emph{Club}=Pressure, \emph{Diamond}=Leverage. These power journeys and gates.

\paragraph{Deck of Consequences (scene drama).}
\emph{Hearts}=social fallout, \emph{Spades}=harm/escalation, \emph{Clubs}=material cost, \emph{Diamonds}=magical/spiritual disturbance.

\textit{Guidance:} Never mix suit meanings across decks. When a rule references ``Spade/Club/Diamond,'' it means \emph{Travel}. When it says ``Hearts/Spades/Clubs/Diamonds,'' it means \emph{Consequences}.

Each generator follows a consistent pattern: draw until you have the required suits, let the highest rank determine the clock size, and use the specific suit meanings to inspire coherent narrative elements. This system ensures that random draws create meaningful complications and opportunities rather than disconnected fragments.

\section{Rank Severity and Clock Size}

The card rank determines the size of the primary Clock for the scene or mission:

\begin{itemize}
\item 2--5 (Minor): 4-segment Clock
\item 6--10 (Standard): 6-segment Clock
\item J, Q, K (Major): 8-segment Clock
\item Ace (Pivotal): 10-segment Clock
\end{itemize}

Color influences tone:
\begin{itemize}
\item Black suits (\textcolor{black}{$\spadesuit$}, \textcolor{black}{$\clubsuit$}): travel hazards, tangible threats, fatigue
\item Red suits (\textcolor{red}{$\heartsuit$}, \textcolor{red}{$\diamondsuit$}): social intrigue, reputational pressure
\end{itemize}

\section{Draw Procedures}

\subsection{Quick Hook (2 cards)}
Draw one Spade and one Heart. The Spade provides the place, the Heart the faction. Use the higher rank to set the Clock.

\subsection{Full Seed (4 cards)}
Draw until one card of each suit appears:
\begin{enumerate}
\item Spade = location
\item Heart = main actor/faction
\item Club = complication
\item Diamond = reward/leverage
\end{enumerate}

The highest rank sets the main Clock. If multiple face cards or Aces appear, begin parallel Clocks.

\subsection{Act Builder}
For each act or session, draw three cards: setting, actor, complication. Save Diamonds to foreshadow leverage or as act payoffs.

\section{Using the Deck}

\begin{enumerate}
  \item Player rolls; each 1 generates a Complication Point (CP).
  \item GM chooses one method for that roll:
  \begin{enumerate}
    \item \textbf{Direct Spend}: translate CP into immediate consequences/clock ticks; or
    \item \textbf{Deck Draw}: draw up to \textbf{min(CP, 3)} cards and \textbf{synthesize a single twist}
    guided by suit and highest rank.
  \end{enumerate}
\end{enumerate}

\section{Combo Rules}

\begin{description}
\item[Pair (same rank)] Recurring motif with a twist
\item[Run (3+ sequential ranks)] Momentum---reduce the main Clock by 1 segment
\item[Flush (3+ same suit)] Strongly theme the act toward that axis
\item[Face + Ace] Reveal a hidden patron or power behind the drawn element
\item[All one color] GM gains 1 free Complication Point in that scene
\end{description}

\section{Regional Generator Summary}

\begin{table}[htbp]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Region} & \textbf{Theme} & \textbf{Special Mechanics} \\
\midrule
Acasia & Broken Marches & Curse mechanics; every A adds lingering omen \\
Aelaerem & Hearth \& Hollow & Red thread motifs; A echoes quiet bells/watch-geese \\
Aeler & Crowns \& Under-Vaults & Stone/breath motifs; A keys click, bells answer \\
Aelinnel & Stone, Bough, Bright Things & Moonlight motifs; A adds shortcut where none should be \\
Black Banners & Condotta \& Crowns & War \& winter motifs; A weapons remember, ice holds dead \\
Ecktoria & Marble \& Fire & Imperial forms; A carves precedent in marble \\
Kahfagia & Pilot's Mirror & Lantern-law jurisdiction shifts; A redefines lanes \\
Linn & Skerries \& Storm-Oaths & Sea omens; A horns on wind, white horses on swell \\
Mistlands & Bells, Salt, Breath & Breath/boundary motifs; A bells answer across water \\
Silkstrand & City of Bridges \& Dyewater & Acasia-is-cursed vibes; A adds lingering omen \\
Theona & Three Greens, No Ninth & ``No Ninth'' custom; A adds omission \\
Thepyrgos & City of a Thousand Stairs & Height/sound motifs; A echoes bells/wind/stair-echo \\
Ubral & Stone Between Spears & Upland motifs; A echoes horns/heather/stone \\
Valewood & Empire Under Leaves & Empire echoes (J/Q/K add relic-logic); A actively rearranges \\
Vhasia & Fractured Sun & Broken-sun motifs; A blots medal/scratches milestone \\
Vilikari & Laurels \& Longhouses & Two-laws motifs; A shows wolf/eagle side-by-side \\
Viterra & Last Kingdom & & \\
Wilds & Roads, Ruins, Weather & Reskin palette for any biome \\
Zakov & Salt \& Serpent & Salt \& serpent motifs; A tides remember, reefs shift, deep listens \\
\bottomrule
\end{tabular}
\caption{Regional Generator Summary}
\end{table}

\input{generators/acasia}
\input{generators/aelaerem}
\input{generators/aeler}
\input{generators/aelinnel}
\input{generators/black_banner}
\input{generators/ecktoria}
\input{generators/kahfagia}
\input{generators/linns}
\input{generators/mistlands}
\input{generators/silkstrand}
\input{generators/theona}
\input{generators/thepyrgos}
\input{generators/ubral}
\input{generators/valewood}
\input{generators/vhasia}
\input{generators/vilikari}
\input{generators/viterra}
\input{generators/wilderness}
\input{generators/ykrul}
\input{generators/zakov}

\section{NPC Generation Deck}

In Fate's Edge, every character should feel like a living, breathing person with their own desires, convictions, and contradictions. This NPC Generation Deck provides a quick yet meaningful way to create compelling non-player characters that will populate your world with depth and intrigue.

The deck is divided into four core categories that define the essential aspects of any character's personality and motivation. Each category offers twelve distinct options, allowing you to draw one card from each suit to rapidly assemble a complete character profile. The combinations are designed to spark narrative possibilities and create characters who feel authentic rather than archetypal.

\textbf{Ambition} defines what the character actively seeks to achieve or obtain. This is their driving goal, the thing that motivates their current actions and decisions.

\textbf{Belief} represents the core principle or philosophy that guides their worldview. This is what they hold as fundamentally true about the world and their place in it.

\textbf{Attitude} describes how they present themselves and interact with others in their day-to-day behavior. This is their social mask and default approach to relationships.

\textbf{Twist} provides an element that contradicts or complicates the other three traits, creating internal conflict and narrative tension. This is where you'll find the character's hidden depths, unexpected vulnerabilities, or surprising contradictions that make them memorable.

When generating an NPC, draw one card from each category and consider how these elements interact with each other. The tension between a character's public ambition and private beliefs, or the contrast between their surface attitude and hidden twist, will naturally suggest story hooks, potential conflicts, and opportunities for character development. This system ensures that even minor NPCs have the potential for complexity and narrative significance.

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
\textbf{Ambition} & \textbf{Belief} & \textbf{Attitude} & \textbf{Twist} \\
\midrule
Power & Might makes right & Arrogant & Secretly insecure \\
Wealth & Ends justify means & Charismatic & Betraying their allies \\
Revenge & Honor above all & Cold & Working for their enemy \\
Love & Truth is sacred & Friendly & Hiding a dark past \\
Knowledge & Loyalty is paramount & Paranoid & Actually an imposter \\
Survival & Family above all & Cruel & Deeply compassionate \\
Fame & Justice must prevail & Pious & Corrupted by power \\
Freedom & Fate can be changed & Optimistic & Hopelessly cynical \\
Protection & Tradition must be upheld & Pessimistic & Revolutionary at heart \\
Control & Change is necessary & Calculating & Acting on impulse \\
Recognition & The system works & Naive & Cynical manipulator \\
\bottomrule
\end{tabular}
\caption{NPC Generation Deck - Core Traits}
\end{table}

\end{chapter}

