\chapter{Travel Reference}

\section{Core Travel Procedure}\index{Travel!Core Procedure}

For each leg of a journey, draw 3--4 cards using the decks for your destination and controlling authority.

\begin{itemize}
\item Spade from the destination deck: sets the scene (place).\index{Travel!Spade}
\item Heart from the destination deck: introduces the local actor or faction.\index{Travel!Heart}
\item Club from the Wilds (general hazards) or destination (if strongly policed): brings pressure.\index{Travel!Club}
\item Diamond from the authority that gates the route: papers, escorts, rights, or exceptions.\index{Travel!Diamond}
\end{itemize}

Set a travel clock by the highest rank:
\begin{itemize}
\item 2--5 $\Rightarrow$ 4 segments\index{Travel!Clocks!4-segment}
\item 6--10 $\Rightarrow$ 6 segments\index{Travel!Clocks!6-segment}
\item J/Q/K $\Rightarrow$ 8 segments\index{Travel!Clocks!8-segment}
\item A $\Rightarrow$ 10 segments\index{Travel!Clocks!10-segment}
\end{itemize}

On success, advance to the next leg; on failure, mark delay, debt, or diversion and resolve a consequence in the fiction.\index{Travel!Consequences}

\section{Mode Rules}\index{Travel!Mode Rules}

\subsection{Sea Legs}\index{Travel!Sea Legs}
(Amaranthine/Dolmis/Aberderrin). If Theona or Valewood 9s show up anywhere in the seed, add an omission or taboo to the leg (a missing step, an unsaid name).\index{Travel!Taboos}

\subsection{Passes Underways}\index{Travel!Underways}
(Aeler). Any A may convert a surface route to an under-route. Diamond from Aeler always codifies outcomes---no roll---for sealed doors, breath time, and escorted segments.\index{Aeler!Underways}

\subsection{Rivers}\index{Travel!Rivers}
Bridges, booms, and law in Ecktoria/Viterra; reed-mazes and bell-lines in Mistlands/Linn waters.\index{Rivers}

\subsection{Frontier Blends}\index{Travel!Frontier Blends}
When origin and destination disagree on law, draw two Diamonds (one from each law) and choose which you will be judged by at the end of the leg.\index{Travel!Law Boundaries}

\section{Regional Routes}\index{Travel!Regional Routes}

\subsection{Amaranthine Coastway}\index{Travel!Amaranthine Coastway}
Kahfagia $\rightarrow$ Ecktoria $\rightarrow$ Acasia $\rightarrow$ Marcott (Vhasia) $\rightarrow$ Fairport (Viterra).

\begin{itemize}
\item Draw Spade+Heart from the port you are entering.
\item Draw Club from the Wilds (gales, quarantine, harbor booms).
\item Draw Diamond from the port's deck (permits, berths), or Kahfagia if convoy/lantern law applies.
\end{itemize}

Note: Silkstrand is Acasia's sole metropolis and a major trading hub; even land caravans risk the marches to reach it.\index{Acasia!Silkstrand}

\subsection{Astroegro Straits}\index{Travel!Astroegro Straits}
Thepyrgos controls the hinge between seas.\index{Thepyrgos!Straits}

\begin{itemize}
\item Draw Spade+Heart and Club from Thepyrgos (stairs, boom, synod).
\item Draw Diamond from Thepyrgos (harbor pass, crane allotment).
\item Switch to Dolmis/Amaranthine modules once the boom is lowered.
\end{itemize}

\subsection{Dolmis Circuits}\index{Travel!Dolmis Circuits}
Fairport (Viterra) $\rightarrow$ Theona (Three Greens) $\rightarrow$ Ubral fjords $\rightarrow$ Aelinnel west shore.

\begin{itemize}
\item Spade+Heart from destination; Club from Linn or the Wilds; Diamond from Theona when island custom matters.\index{Theona!Dolmis}
\end{itemize}

\subsection{Aelerian Passes Underways}\index{Travel!Aelerian Passes}
Vhasia/Viterra/Ubral south slopes $\rightarrow$ Aeler gates $\rightarrow$ Mistlands.

\begin{itemize}
\item Spade+Heart and Club from Aeler; Diamond from Aeler (Underway Pass, Key-Writ, Breath Time).\index{Aeler!Gates}
\end{itemize}

\subsection{Shadow Corridors}\index{Travel!Shadow Corridors}
Thin Shore (Valewood east coast): risky misted corridor north--south toward Zakov.\index{Valewood!Shadow Corridors}

\begin{itemize}
\item Spade from Valewood, Heart from Mistlands (or Valewood), Club from Mistlands (wraiths, bell-line failures), Diamond from either Mistlands (Ward-salt, Lantern Writ) or Valewood (Way-cord, Truce-bough) depending on which law you invoke.
\item Rule of 9s: any 9 in the seed adds an omission.\index{Travel!Rule of 9s}
\end{itemize}

Green Gate hops (Aelinnel): tide-timed arches.\index{Aelinnel!Green Gate}

\begin{itemize}
\item Spade+Club from Aelinnel (tides, geasa), Diamond from Aelinnel (Tide-path Key, Host-rights).\index{Tides}
\end{itemize}

\subsection{River Roads}\index{Travel!River Roads}
Belworth: forms the boundary between Vhasia and Viterra. Twin ports at the mouth (Marcott and Fairport); at mid-reach Tarlington Falls requires lift/portage up to the plateau town. The Lence joins here from Vhasia. Use Viterra for east-bank audits/bridges, Vhasia for west-bank politics.\index{Belworth River}

Yloka: raids run down from the north into the Aberderrin and Dolmis---overlay Linn Clubs and draw destination for Spade+Heart.\index{Yloka}

Acasian rivers: hazardous but profitable approaches to Silkstrand; use Acasia Clubs (feuds/levies/curse) and Diamonds (condotta, bridge rights).\index{Acasia!Rivers}

\subsection{Steppe Frontiers (Violet Steppes \& Meadows)}\index{Travel!Steppe Frontiers}
Ykrul $\leftrightarrow$ Vilikari $\leftrightarrow$ Ecktoria/Acasia borders.\index{Ykrul!Frontiers}\index{Vilikari!Frontiers}

\begin{itemize}
\item Spade from the road you ride (Wolf Road or Foedus Stone), Heart from the counterpart (envoys, comitatus), Club from Wilds or frontier decks (rasputitsa, hostage protocol, foedus recall), Diamond from either Ykrul (paiza, remounts) or Vilikari (foedus seal, bench-right).\index{Foedus Stone}
\end{itemize}

\section{Gateways Control Points}\index{Travel!Gateways}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
\textbf{Gate} & \textbf{Deck} & \textbf{Diamond Source} & \textbf{Typical Leverage} \\
\midrule
Kassamira (Kahfagia) & Kahfagia & Kahfagia & Convoy letter; lantern-law warrant; pilotage token \\
Ecktoria Capital Port & Ecktoria & Ecktoria & Berth priority; coin-house credit; procession permit \\
Silkstrand & Acasia & Acasia & Exchange floor; condotta; watergate priority \\
Marcott & Vhasia & Vhasia & Letters patent; safeconduct; fairground precedence \\
Fairport & Viterra & Viterra & Customs seal; ferry priority; Dawn escort \\
Tarlington (plateau lift) Falls & Viterra & Viterra & Lift priority; bridge/portage pass; County Thing ruling \\
Thepyrgos (boom) & Thepyrgos & Thepyrgos & Harbor pass; crane allotment; Synod indulgence \\
High-Mist Pass & Aeler & Aeler & Underway pass; Key-Writ; Breath time allotment \\
Payden's Port & Mistlands & Mistlands & Lantern writ; ward-salt; wraith indemnity \\
Thin Shore (Valewood) & Valewood & Valewood/Mistlands & Way-cord; dew-mirror; lantern writ \\
Green Gate (Aelinnel) & Aelinnel & Aelinnel & Tide-path key; host-right bough; salt license \\
Ubral Passes & Ubral & Aeler/Ubral & Vurim pass-ring; hill-fort shelter writ; guide's braid \\
Thing-holm (Linn) & Linn & Linn & Thing ruling; harbor-green; escort writ \\
Foedus Stone (March) & Vilikari & Vilikari/Ykrul & Foedus renewal; remount allotment; hostage treaty \\
Zakov Roadstead & Theona/Linn & Theona/Linn & Raid-truce ribbon; whale-road escort; market immunity \\
\bottomrule
\end{tabular}
\end{table}

\section{Special Rules \& Taboos}\index{Travel!Special Rules}

\subsection{Theona Valewood 9s}\index{Travel!Taboos}
Whenever a 9 appears in a seed involving Theona or Valewood, add an omission: a missing step, an unsaid name, an unseen guest. If the 9 is a Diamond, you may break the taboo once---someone will come to collect.\index{Theona!Taboos}\index{Valewood!Taboos}

\subsection{Aeler Aces}\index{Travel!Aces}
An A can rewire routes beneath the mountains. You may swap your Spade to Aeler (keeping the other seed cards) and continue under-vault procedures.\index{Aeler!Aces}

\subsection{Echoing Omens}\index{Travel!Echoing Omens}
Any A adds a motif you can echo later (ash on the wind, a bell heard underground, a landmark that returns elsewhere). Make the world feel like it remembers.\index{Omens}

\section{Worked Itineraries}\index{Travel!Worked Examples}

\subsection{West-to-East Coastal Haul (Kahfagia $\rightarrow$ Viterra)}\index{Travel!Coastal Haul}

\textbf{Leg 1: Kassamira $\rightarrow$ Ecktoria}

Seed: Ecktoria Spade (Aqueduct arcades), Ecktoria Heart (Coin-house factor), Wilds Club (gale), Kahfagia Diamond (convoy letter).

Clock: 6. Read: cross under coin-house eyes with convoy priority; bad weather strains schedules.

\textbf{Leg 2: Ecktoria $\rightarrow$ Silkstrand}\index{Silkstrand!Approach}

Seed: Acasia Spade (Three-Queens Bridge), Acasia Heart (Dyers' Guildmistress), Acasia Club (loom strike), Acasia Diamond (Exchange pass).

Clock: 6--8 depending on unrest.

\textbf{Leg 3: Silkstrand $\rightarrow$ Marcott}\index{Marcott!Approach}

Seed: Vhasia Spade (Pont-du-Tithe), Vhasia Heart (Parlement clerk), Vhasia Club (coin rumor), Vhasia Diamond (letters patent).

\textbf{Leg 4: Marcott $\rightarrow$ Fairport}\index{Fairport!Approach}

Seed: Viterra Spade (Fairport tideworks), Viterra Heart (shipwright), Linn Club (boom lifts), Viterra Diamond (customs seal).

Clock: 6.

\subsection{Acasia $\rightarrow$ Mistlands (Forgotten Pass + Under-Gate)}\index{Travel!Mountain Passes}

\textbf{Leg 1: Silkstrand $\rightarrow$ Aeler Gate}\index{Aeler!Gate Approach}

Seed: Aeler Spade (Avalanche gallery), Aeler Heart (Geometer), Aeler Club (Engineer requisition), Aeler Diamond (Underway Pass).

\textbf{Leg 2: Gate $\rightarrow$ Mistlands}\index{Mistlands!Approach}

Seed: Mistlands Spade (Bell-Line levee), Mistlands Heart (Bell-warden), Mistlands Club (wraith crossing), Mistlands Diamond (Ward-salt).

\subsection{Thin Shore $\rightarrow$ Zakov Corridor}\index{Travel!Shadow Routes}

\textbf{Leg 1: Payden's Port $\rightarrow$ Thin Shore}\index{Thin Shore!Approach}

Seed: Valewood Spade (Green lane), Mistlands Heart (Protectorate clerk), Mistlands Club (bell failure), Mistlands Diamond (Lantern writ).

\textbf{Leg 2: Thin Shore transit}\index{Thin Shore!Transit}

Seed: Valewood Spade (Unfound stile), Valewood Heart (Path-warden), Valewood Club (Sweet wind), Valewood Diamond (Way-cord).

Rule of 9s applies.

\subsection{Frontier Mediation (Ykrul $\leftrightarrow$ Vilikari)}\index{Travel!Frontier Mediation}

\textbf{Leg: Foedus Stone parley}\index{Foedus Stone!Parley}

Seed: Vilikari Spade (Foedus Stone), Ykrul Heart (Noyan envoy), Wilds Club (rasputitsa), Vilikari Diamond (Foedus Renewal).

Blend: also draw a Ykrul Diamond (Paiza). Choose which law saves your schedule.\index{Foedus}

\section{Pace, Clocks, and Consequences}\index{Travel!Pace}

\begin{itemize}
\item Road/river leg: 6-segment clock; +2 segments if crossing a law boundary (new passports, new prayers).\index{Travel!Law Boundaries}
\item Sea strait/mountain pass: 8-segment clock; --2 segments if you spend a Diamond that codifies priority (pilotage, pass, convoy).\index{Travel!Priority}
\item Shadow corridor: 6-segment clock; on any A, add a free omen to echo later.\index{Travel!Shadow Corridors}
\item Failure defaults: delay (lose time; an enemy advances), diversion (alternate route; new Club), or debt (promise, fine, or favor owed to the authority that controls the gate).\index{Travel!Failure Consequences}
\end{itemize}
\end{chapter}
