\chapter{Resource Management}

\section{Supply Clock}

The Supply Clock is a shared condition for the entire party, representing food, water, and basic gear.

\subsection{States}

\begin{description}
\item[Full Supply (0 filled)] The party is well-equipped.
\item[Low Supply (2 filled)] Minor narrative complications: bland food, damaged arrows, thinning waterskins.
\item[Dangerously Low (3 filled)] Each character gains Fatigue.
\item[Out of Supply (4 filled)] Severe penalties; starvation, dehydration, failing gear.
\end{description}

\subsection{Filling the Clock}

\begin{itemize}
\item Harsh travel or lost pack animals (GM fiat)
\item GM spends 2+ Complication Points
\item The party chooses to travel light for advantage
\end{itemize}

\subsection{Emptying the Clock}

\begin{itemize}
\item Reaching civilization resets to Full
\item Foraging/hunting: group Survival check clears 1 segment
\item Downtime in safety removes 1 segment
\end{itemize}

\section{Fatigue}

Fatigue represents exhaustion, hunger, and strain.

\subsection{Effects}

\begin{description}
\item[1 Fatigue] Minor drain. On your next roll, re-roll one success (player's choice).
\item[2 Fatigue] Worn down. On each roll, re-roll one success.
\item[3 Fatigue] Failing fast. On each roll, re-roll two successes.
\item[4 Fatigue] Collapse/KO/Spiritual break. You fall out of the scene until treated or rescued.
\end{description}

\subsection{Clearing Fatigue}

A night's rest with adequate Supply removes 1 level. You cannot clear Fatigue if the party's Supply clock is Dangerously Low or Empty.

\section{Gear Condition}

Gear does not have hit points. It suffers only when drama demands it.

\subsection{Compromised Items}

\begin{itemize}
\item Introduced via Complication Points or narrative consequence
\item A Compromised item gives --1 die on relevant rolls
\end{itemize}

\subsection{Breaking Point}

If a Compromised item suffers another setback, it breaks entirely.

\subsection{Repair}

\begin{description}
\item[Field Repair] Temporary; Craft or Survival check removes penalty for one scene
\item[Proper Repair] Permanent; requires tools, materials, and downtime
\end{description}

\section{Asset and Follower Management}

\subsection{Followers (On-Screen)}

\begin{description}
\item[Cost] Cap$^2$ XP
\item[Assist Dice] When applicable, the follower adds help dice equal to \textbf{min(C, the helper's relevant Skill)}, capped at +3 dice. Exception: The "Exceptional Coordination" Talent allows one follower to provide +4 assist dice.
\item[Cap] 1-5 (5 is exceptional)
\item[Upkeep] Each Downtime, pay XP equal to C or spend a Scene tending the relationship.
\item[Risk] If the GM spends 2+ Complication Points on an action you take with assistance, they may endanger, injure, or separate the follower instead of you if fictionally appropriate.
\item[Off-Screen Capability] Once per downtime, a Cap 5 follower can solve one significant problem but generates 1 CP for crew. The GM must describe how the follower's action creates story consequences for the CP generated.
\end{description}

\subsection{Follower Assist}
\begin{itemize}
    \item Assist dice come from the helper, not the leader.
    \item Total Assist on any roll (from any sources) remains hard-capped at +3. Exception: The "Exceptional Coordination" Talent allows one follower to provide +4 assist dice.
    \item Slot Limit: Only one follower may assist a given action.
\end{itemize}

\subsection{Loyalty \& Bonds (Optional)}
\begin{itemize}
    \item Track a simple Loyalty tag per follower: Wary / Steady / Devoted.
    \item Devoted followers can once per arc convert one GM Complication targeting them into a lesser setback.
    \item Wary followers cost +1 XP to Maintain.
\end{itemize}

\subsection{Stress, Harm, \& Loss (GM Tools)}
\begin{itemize}
    \item \textbf{Pin}: The follower is separated/boxed out.
    \item \textbf{Wound}: The follower is Injured: until treated off-screen, their Cap counts as 1 lower.
    \item \textbf{Burn}: Mark Neglected immediately.
    \item \textbf{Seize}: Escalate to Compromised.
    \item \textbf{PC Choice Lever}: The GM should offer the player a save.
\end{itemize}

\subsection{Off-Screen Assets}

\begin{description}
\item[Minor (4 XP)] Safehouse, small charter
\item[Standard (8 XP)] Noble title, guild section, spy ring
\item[Major (12 XP)] City license, regional network
\item[Activation] 1 Boon
\end{description}

\subsection{Asset Activations}
\begin{itemize}
    \item Off-Screen effects: Use each Asset's listed Off-Screen effect once per session for free.
    \item On-Screen activations: To reshape the current scene, spend 1 Boon.
    \item Plausibility test: The Asset must have scope and reach.
\end{itemize}

\subsection{Condition Tracks}

\begin{description}
\item[Maintained] Full capability; no penalties
\item[Neglected] --1 die when used (assist or leverage). Narratively: slower, sullen, short-staffed
\item[Compromised] Unavailable. Narratively: captured, burned, seized, defected
\end{description}

\subsection{Maintenance and Repair}

\begin{itemize}
\item After a story arc or 2--3 sessions of heavy use, the GM may flag the resource as at risk
\item To keep it Maintained, the player must either:
  \begin{itemize}
  \item Spend downtime repairing/servicing it, or
  \item Pay XP immediately (representing replacement parts, masterwork upkeep)
  \item If neglected, the gear becomes Compromised (--1 die penalty) and may eventually break
  \item Superior and Artifact gear do not require upkeep, but if Compromised through Complications, only narrative quests can repair them
  \end{itemize}
\end{itemize}

\subsection{Bond-Driven Resource Generation}
When a player takes a significant action to aid an ally with whom they share a bond, and explicitly references that bond in an intricate description, they may mark that bond to gain 1 boon after the action resolves.

\textbf{Requirements:}
\begin{itemize}
    \item \textbf{Mutual Bond:} Player shares a bond with the ally they're aiding
    \item \textbf{Intricate Description:} Player describes how their bond motivates their action
    \item \textbf{Significant Aid:} Meaningful assistance (not just +1 die)
    \item \textbf{Fiction First:} The bond must genuinely drive the choice to help
\end{itemize}

\textbf{Examples:}
\begin{itemize}
    \item "Remembering how they saved me before, I throw myself in front of the attack meant for them!"
    \item "Thinking of our shared past, I use my last resource to shield them from harm."
    \item "Drawing on our mutual experiences, I rally others to keep fighting alongside them."
\end{itemize}

\textbf{Limitations:}
\begin{itemize}
    \item Once per bond per session
    \item Must be a meaningful sacrifice or risk
    \item GM approval for "significant action"
    \item Cannot be used for basic assistance rolls
\end{itemize}

\section{Over-Stack}
\begin{itemize}
    \item Structural advantages: active buff/tag, favorable venue/pennant, Follower Initiative unused, on-screen Asset activation, opponent disadvantaged by fiction, ritual prep that applies now.
    \item Trigger: If the crew enters a scene with ≥ 3 structural advantages, apply Over-Stack once for that scene: either start one named rail at +1 or the GM banks +1 CP for the first Deck Twist.
\end{itemize}
\end{chapter}
