\chapter{Magic System}

\section{Design Philosophy}\index{Magic!Design Philosophy}

Magic in Fate's Edge is not a tool of convenience but a dangerous negotiation with the fabric of reality. It is powerful, flexible, and transformative---yet every attempt to shape it carries risk. The dice never merely ask "does it work?" but always whisper "what is the cost?"

\section{The Nature of Magic}\index{Magic!Nature}

\begin{itemize}
\item \textbf{Volatile by Design}: Magic is not fully understood, even by its most adept practitioners. Every working pushes against boundaries that resist being bent.\index{Magic!Volatility}
\item \textbf{Risk Embodied}: Each spell generates Complication Points. These points do not vanish; they manifest as Backlash, unpredictable consequences that ripple outward.\index{Magic!Risk}
\item \textbf{Narrative Weight}: Casting is always a story moment. Even a "successful" spell alters the scene in ways the caster did not intend.\index{Magic!Narrative Weight}
\item \textbf{Thematic Consequence}: Backlash is not arbitrary; it aligns with the opposing or uncontrolled element of the Art invoked (flame flares out of control, shadows linger too long, storms roll beyond command).\index{Magic!Thematic Consequence}
\end{itemize}

\section{The Eight Elements of Magic}\index{Magic!Elements}

The fundamental metaphysical system that underlies all magical practice in Fate's Edge divides magic into eight distinct elements, each with its own nature and resonance.

\subsection{Physical Elements}\index{Magic!Elements!Physical}

\subsubsection{Earth}\index{Magic!Elements!Earth}
\textbf{Nature:} Solidity, stability, foundation\\
\textbf{Domain:} Stone, soil, metal, structure\\
\textbf{Manifestation:} Shaping, sensing, moving earth and stone\\
\textbf{Backlash Theme:} Rigidity, collapse, awakening

\subsubsection{Fire}\index{Magic!Elements!Fire}
\textbf{Nature:} Energy, transformation, destruction\\
\textbf{Domain:} Flame, heat, light, change\\
\textbf{Manifestation:} Ignition, heating, burning, purifying\\
\textbf{Backlash Theme:} Uncontrolled spread, scorching, consuming

\subsubsection{Air}\index{Magic!Elements!Air}
\textbf{Nature:} Movement, speed, freedom\\
\textbf{Domain:} Breath, sound, motion, flight\\
\textbf{Manifestation:} Moving objects, creating drafts, amplifying sound\\
\textbf{Backlash Theme:} Whipping, dispersal, loss of control

\subsubsection{Water}\index{Magic!Elements!Water}
\textbf{Nature:} Fluidity, healing, adaptability\\
\textbf{Domain:} Liquids, tides, rain, flow\\
\textbf{Manifestation:} Calming, directing, purifying water\\
\textbf{Backlash Theme:} Stagnation, flooding, contamination

\subsection{Metaphysical Elements}\index{Magic!Elements!Metaphysical}

\subsubsection{Fate}\index{Magic!Elements!Fate}
\textbf{Nature:} Destiny, inevitability, long-term consequences\\
\textbf{Domain:} Time, prophecy, causality, \emph{anti-magic}\\
\textbf{Manifestation:} Seeing future threads, influencing destiny, sealing oaths\\
\textbf{Backlash Theme:} Paradox, temporal loops, unintended consequences

\subsubsection{Life}\index{Magic!Elements!Life}
\textbf{Nature:} Vitality, creation, spark of existence\\
\textbf{Domain:} Health, growth, repair, birth\\
\textbf{Manifestation:} Healing, growth, vitality enhancement\\
\textbf{Backlash Theme:} Overgrowth, fever, untidy healing

\subsubsection{Luck (Fortune)}\index{Magic!Elements!Luck}
\textbf{Nature:} Chance, unpredictability, probability\\
\textbf{Domain:} Randomness, breaks, windfalls\\
\textbf{Manifestation:} Influencing chance, creating opportunities\\
\textbf{Backlash Theme:} Side-coincidence, misfire elsewhere, ironic reversals

\subsubsection{Death/Dreams (Obishaal)}\index{Magic!Elements!Death}
\textbf{Nature:} Endings, thresholds, subconscious, \emph{Ways Between}\\
\textbf{Domain:} Sleep, death, spirits, doorways\\
\textbf{Manifestation:} Communication with the dead, dream manipulation, passage\\
\textbf{Backlash Theme:} Hauntings, nightmares, thin walls, threshold opens

\section{Magical Arts and Personal Expression}\index{Magic!Arts}

Your \emph{Art} is the narrative method by which your \textbf{Weave} and \textbf{Cast} work (sigils, sung names, lantern-law, bone charms, contracts, salt-thread, etc.).

\subsection{Declaring an Art}\index{Magic!Arts!Declaring}
When you gain \textbf{Arcane Initiate}, write a one–two line \textbf{Art} describing:
\begin{itemize}
  \item \textbf{Gesture \& medium} (ink, chord, breath, light, bone, law).
  \item \textbf{Typical Elements} you lean on (pick 2 you're often aligned to).
\end{itemize}

\subsection{Art in Play}\index{Magic!Arts!In Play}
\begin{itemize}
  \item \textbf{Spotlight bump (1/scene):} if your Art is \emph{clearly honored in fiction} (right tools/time/setting), gain \textbf{+1 die} on your \textbf{Cast}. This counts toward the +3 dice cap.
  \item \textbf{Off-style strain:} if you're forced to work wildly \emph{against} your Art (no tools, hostile locus), the GM may set a worse \textbf{Position} or require you to accept extra \textbf{Backlash} choices on a Partial.
\end{itemize}

\subsection{Examples}\index{Magic!Arts!Examples}
\begin{itemize}
  \item \emph{Sealwright's Chant} — lantern, chain, chalk; Elements: Earth/Fire; \textit{I sing hinges into hearing.}
  \item \emph{Salt-Thread Scribe} — salt, silk, breath; Elements: Water/Luck; \textit{Knots remember routes and debts.}
  \item \emph{Name-Caller} — true names, oaths, bells; Elements: Fate/Air; \textit{Words stand up when I say them.}
\end{itemize}

\section{Casting Procedure}\index{Magic!Casting Procedure}

All spellcasting follows a structured sequence called the Casting Loop. It unfolds across two phases of play: gathering strength, then weaving it into form.

\subsection{Channel}\index{Magic!Channel}
The caster focuses, rolling Wits + Arcana to gather Potential. Each success becomes fuel for shaping the spell. Each 1 adds Complication Points immediately.

\subsection{Weave}\index{Magic!Weave}
On the following turn, the caster rolls Wits + (Art) to shape Potential into a defined effect. The Description Ladder applies: Basic/Detailed/Intricate descriptions reduce or redirect Complication Points.

\subsection{Backlash}\index{Magic!Backlash}
Complication Points spent by the GM manifest as uncontrolled consequences. These are thematic to the Art and scale with the number of points spent: minor nuisances at low levels, dangerous disasters at high levels.

\textbf{Mitigation}: Boons do not reduce CP unless a Talent/Asset explicitly says "Mitigate CP."

\section{Freeform Spellcasting}\index{Magic!Freeform Casting}

Use this flow for improvised magic:

\begin{enumerate}
  \item \textbf{Declare}: Describe your Weave and name 1–2 Elements that color it.
  \item \textbf{Set DV}: GM sets by scope and situation (see DV Guidance table).
  \item \textbf{Roll Weave}: Wits + Art; 6+ = success, 1 = CP.
  \item \textbf{Apply Outcome}: Use standard Outcome Matrix.
  \item \textbf{Backlash}: On Partial/Miss (or two 1s on a Hit), color the cost by chosen Element or its opposite.
\end{enumerate}

\subsection{DV Guidance}\index{Magic!DV Guidance}
\begin{tabular}{@{}llp{9cm}@{}}
\textbf{DV 2} & Small/local & One target, Near band, short veil, fragile barrier, simple shove/cleanse. Aligned Realm/tools may bump Effect to \emph{Standard}. \\
\textbf{DV 3} & Scene-scale & Small zone multi-target, sturdy barrier [2/4], move one band, strong veil/reveal, oath ping. Opposed Realm may drop Effect to \emph{Limited}. \\
\textbf{DV 4} & Big swing & Zone control, Far reach, complex transport, high-quality ward, reshape Position broadly. Expect backlash if rushed. \\
\textbf{DV 5+} & Set-piece/ritual & Battlefield rewriting, clause-level edits of reality, defying entrenched wards. Wants prep, aids, or locus. \\
\end{tabular}

\subsection{Backlash by Element}\index{Magic!Backlash!Elemental}
\begin{itemize}
  \item \textbf{Earth} $\rightarrow$ rubble, pin, heavy footing; \textbf{vs Air} $\rightarrow$ sound carries, you're exposed.
  \item \textbf{Fire} $\rightarrow$ burns, flares, unwanted attention; \textbf{vs Water} $\rightarrow$ slick, shorted gear, sodden scrolls.
  \item \textbf{Air} $\rightarrow$ scatter, drop, vertigo; \textbf{vs Earth} $\rightarrow$ stuck, cramped, dust choke.
  \item \textbf{Water} $\rightarrow$ leak, flood, cold drag; \textbf{vs Fire} $\rightarrow$ smoke, sputter, dim.
  \item \textbf{Fate} $\rightarrow$ options close, "only one way"; \textbf{vs Luck} $\rightarrow$ mischance hits an ally.
  \item \textbf{Luck} $\rightarrow$ side-effect elsewhere, fragile success; \textbf{vs Fate} $\rightarrow$ your gambit snaps to a fixed, harsher outcome.
  \item \textbf{Life} $\rightarrow$ overgrowth, fever, hungry repair; \textbf{vs Death/Dreams} $\rightarrow$ numbness, sleep-tug, a door opens.
  \item \textbf{Death/Dreams} $\rightarrow$ fade, echo, threshold opens/closes; \textbf{vs Life} $\rightarrow$ pain returns, rot sets in, breath drags.
\end{itemize}

\section{Ritual Casting (Optional Rule)}\index{Magic!Ritual Casting}
\begin{itemize}
    \item \textbf{Ritual Helper Cap}: You may draw on ceil(Arcana/2) helpers (max 3).\index{Magic!Ritual!Helpers}
    \item \textbf{Procedure}:
    \begin{enumerate}
        \item Declare the Ritual.\index{Magic!Ritual!Declare}
        \item Channel Together.\index{Magic!Ritual!Channel}
        \item Weave.\index{Magic!Ritual!Weave}
        \item Backlash.\index{Magic!Ritual!Backlash}
    \end{enumerate}
    \item \textbf{Relevant skills}: Helpers may use different relevant skills if their procedure is fictionally distinct.\index{Magic!Ritual!Skills}
    \item \textbf{CP distribution}: CP from Channel resolves on that roller. CP from Weave is assigned to the primary caster.\index{Magic!Ritual!CP Distribution}
\end{itemize}

\section{Rites (Pact Magic)}\index{Magic!Rites}

Rites are precise, named effects granted by powerful Patrons. This framework keeps play fast and consistent.

\subsection{Becoming an Acolyte}\index{Magic!Rites!Becoming}
\begin{itemize}
  \item Choose \textbf{one} or \textbf{two Patrons}. (Two is allowed; see \emph{Cross-Patron Interference}.)
  \item Start with \textbf{2 Low} and \textbf{1 Standard} Rite total (split across your Patrons as you wish).
  \item Track an \textbf{Obligation ledger} \emph{per Patron} (segments of debt accrued through use).
\end{itemize}

\subsection{Using a Rite}\index{Magic!Rites!Using}
Unless a Rite says otherwise:
\begin{enumerate}
  \item \textbf{Invoke (1 action).} Speak the name, draw the sign, or employ the proper tool. If a roll is required, the GM sets \textbf{DV by fiction} (typical scene DV 2–4).
  \item \textbf{Mark Obligation.} Invoking a Rite usually marks \textbf{+1 segment} to that Patron's ledger (some Low rites may be free at GM discretion).
  \item \textbf{Push It (optional).} Amplify the effect; mark \textbf{+1 Obligation}. A given Rite may be Pushed at most \textbf{once per scene}.
  \item \textbf{Backlash (on a 1 or Miss).} The GM inflicts a fitting consequence \emph{or} marks \textbf{+1 Obligation}.
\end{enumerate}

\subsection{Stacking \& Caps}\index{Magic!Rites!Stacking}
\begin{itemize}
  \item \textbf{Same-source scene buffs do not stack.} Take the best version; others are suppressed.
  \item A \textbf{single Rite} can add at most \textbf{+3 Obligation segments per scene} \emph{from that Rite} (Invoke + Push + possible Backlash).
\end{itemize}

\subsection{Cross-Patron Interference}\index{Magic!Rites!Interference}
If you \textbf{switch Patrons in the same scene} (e.g., after using Gate, you Invoke Ikasha):
\begin{itemize}
  \item Immediately mark \textbf{+1 Obligation} to the second Patron (jealous attention).
  \item The GM may color the next moment with minor interference (audible hum, flicker, bystander notice).
  \item Each further \emph{switch} that scene marks \textbf{+1} again.
\end{itemize}

\subsection{Clearing Obligation}\index{Magic!Rites!Clearing}
Between scenes or in downtime, you may clear \textbf{1–2 segments} per Patron by:
\begin{itemize}
  \item \textbf{Service} fitting the Patron (tend hinges for the Gate, keep sanctioned watch for the Oath, etc.), or
  \item Spending \textbf{1–2 Boons} to represent offerings, favors, or perfect observance.
\end{itemize}
At \textbf{+3/+6/+9} total segments, foreshadow manifestations (omens, auditors, bailiffs) as your table prefers.

\subsection{Tags}\index{Magic!Rites!Tags}
Rites may include SRD tags such as \texttt{[WARD]}, \texttt{[BANISH]}, \texttt{[UNWARD]}. \textbf{Tags only function when printed on a specific Talent, Ability, or as the result of a Spell.} For Outsiders, use the unified DV/Leash rules.

\section{Two Sample Patrons}\index{Magic!Rites!Sample Patrons}

\subsection{The Sealed Gate — Boundaries, Jurisdiction, Closure}\index{Magic!Rites!Gate}
\textit{You write borders into the world and prosecute trespass.}

\paragraph{Low Rites.}
\begin{description}
  \item[Seal the Threshold.] Draw a brief sign across a door/line. Enemies crossing suffer worsened position or stumble. \emph{Invoke:} +1 Obl. \emph{Push:} strengthen the edge (e.g., treat as difficult terrain) \emph{(+1 Obl)}.
  \item[Key's Rebuke.] Snap a spectral hasp at a reaching hand or tool; stagger or disarm for a beat. \emph{Invoke:} +1 Obl.
\end{description}

\paragraph{Standard Rites.}
\begin{description}
  \item[Circle of Denial \texttt{[WARD]} (affects Outsiders).] Mark a ring or arc; Outsiders that \emph{cross} must test \emph{Cross the Ward} (\textbf{DV = Cap}). \emph{Hit:} cross and \textbf{+DV} to Leash; \emph{Partial:} cross and \textbf{+1}; \emph{Miss:} fail to cross this beat. \emph{Invoke:} +1 Obl. \emph{Push:} fortify the circle \emph{(+1 Obl)}.
  \item[Writ of Passage.] Name one path as \emph{permitted}; allies on that route gain improved position/flow. \emph{Invoke:} +1 Obl. \emph{Push:} extend to one extra ally or carry across one obstacle \emph{(+1 Obl)}.
\end{description}

\paragraph{High Rites.}
\begin{description}
  \item[Banishment Knot \texttt{[BANISH]} (Outsider).] Cast out a named Outsider at Near. \emph{DV = Cap}. \emph{Hit:} \textbf{+DV} to Leash (often filling); \emph{Partial:} \textbf{+1}. \emph{Invoke:} +1--2 Obl (table preference). \emph{Push:} strip one tether/anchor \emph{(+1 Obl)}.
\end{description}

\subsection{Ikasha — Shadow, Latent Potential, Penumbra}\index{Magic!Rites!Ikasha}
\textit{You bank what might happen and cash it when the light looks away.}

\paragraph{Low Rites.}
\begin{description}
  \item[Borrowed Shade.] Cool or deepen existing dim; allies acting within gain improved position until violence occurs there. \emph{Invoke:} +1 Obl.
  \item[Mark of Penumbra.] Tag a place/object with a small shadow-glyph; later, blur sound or slip attention at that mark once. \emph{Invoke:} +1 Obl. \emph{Push:} extend the tell-skip to one ally \emph{(+1 Obl)}.
\end{description}

\paragraph{Standard Rites.}
\begin{description}
  \item[Umbra Thread.] Tether two shadows you can see; once this scene, you or a willing ally traverse between them as a single Move. \emph{Invoke:} +1 Obl. \emph{Push:} allow one additional traverse or longer reach \emph{(+1 Obl)}.
  \item[Deferred Answer.] Bank one pointed question/command aimed at you; until you answer, you resist pins and scans more easily. When you cash it, answer honestly or accept a complication. \emph{Invoke:} +1 Obl.
\end{description}

\paragraph{High Rite.}
\begin{description}
  \item[Eclipse Clause \texttt{[WARD]} (affects Outsiders).] For one scene, an area is penumbral; ranged targeting beyond Near is impaired; shadow-to-shadow movement is improved. Outsiders crossing the edge use the unified \texttt{[WARD]} crossing (DV = Cap; \emph{Hit} \textbf{+DV} Leash, \emph{Partial} \textbf{+1}). \emph{Invoke:} +2 Obl. \emph{Push:} harden the border for one beat \emph{(+1 Obl)}.
\end{description}

\section{Pact-Whisperer — Standalone Summoning}\index{Magic!Summoning}

\textit{Call a spirit fast; bind it with a tiny leash; keep play fiction-first.}

\subsection{Core Procedure}\index{Magic!Summoning!Procedure}
\begin{enumerate}
  \item \textbf{Call (1 action):} A spirit manifests at \textbf{Near}. Pick a \textbf{Spirit Template}.
  \item \textbf{Bind (no extra roll):} choose one: spend \textbf{1 Boon} \emph{or} mark \textbf{1 Fatigue}.
  \item \textbf{Leash (lives on the spirit):} set \textbf{Leash = Cap + 2} segments. \\
        Cap 1 $\rightarrow$ [3] \hspace{1em} Cap 3 $\rightarrow$ [5]
  \item \textbf{Tick Leash} (no rolls) when any happen: the spirit \textbf{takes harm}; you \textbf{command against its nature}; you \textbf{split focus} (you take another significant action while it acts on your order); a rival \textbf{contests} it; it \textbf{moves from Close to Far}. Crossing a \texttt{[WARD]} uses the rules under \S\ref{sec:tags}.
  \item \textbf{When the Leash fills:} the spirit \textbf{acts to its nature once}, then \textbf{departs}.
\end{enumerate}

\paragraph{Action economy.} Issuing a meaningful command uses \textbf{your action}. \\
\textbf{Limit:} one active spirit at a time (you may Call again after departure). \\
\textbf{Downtime:} all summons \textbf{depart at downtime} unless an ability explicitly says otherwise.

\paragraph{Boon Finesse}\index{Magic!Summoning!Boon Finesse}
Once per round, you may spend \textbf{1 Boon} to \textbf{clear 1 tick} from your current spirit's Leash. You cannot use this after the Leash has already filled.

\paragraph{Talent gates.}\index{Magic!Summoning!Talents}
\begin{itemize}
  \item \textbf{Lesser Pactwright (Talent):} You may Call spirits of \textbf{Cap 1}.
  \item \textbf{Greater Pactwright (Talent):} You may also Call spirits of \textbf{Cap 3}.
\end{itemize}

\section{Outsiders (Definition \& Caps)}\index{Magic!Outsiders}

\textbf{Outsider:} any being \textbf{not native} to this world/setting—e.g., summoned entities, demons, extraplanar creatures, spirits, some celestials, and similar arrivals-from-beyond.

\paragraph{Cap (for Outsiders).}\index{Magic!Outsiders!Cap}
\begin{itemize}
  \item \textbf{PC-summoned Outsiders:} Cap is set by the summoner's talent (Lesser = \textbf{Cap 1}, Greater = \textbf{Cap 3}).
  \item \textbf{World/NPC Outsiders:} the stat block lists Cap; if absent, the GM assigns: Lesser = 1, Greater = 3, Elder = 5 (rare).
\end{itemize}

\section{Tags for Talents, Abilities, and Spell Results}\index{Magic!Tags}

\textbf{Tags apply only when printed on a specific \emph{Talent}, \emph{Ability}, or as the \emph{result of a Spell}.}
They do nothing on their own. Spells are accessible but \textbf{risky}; Talents/Abilities provide safer, repeatable access.

\subsection{\texttt{[WARD]}}\index{Magic!Tags!WARD}
Creates a magical \textbf{edge/circle/zone} that \textbf{challenges Outsiders} (or other targets if the text explicitly says so).

\paragraph{Crossing a \texttt{[WARD]} (when it affects Outsiders).}
\begin{itemize}
  \item Roll a suitable pool (table choice, e.g., Spirit+Lore, Resolve+Lore, Presence+Command).
  \item \textbf{DV = target's Cap.}
  \item \textbf{On a Hit:} the Outsider \textbf{crosses} and its \textbf{Leash gains +DV segments} (i.e., +Cap).
  \item \textbf{On a Partial:} the Outsider \textbf{crosses} and its \textbf{Leash gains +1 segment}.
  \item \textbf{On a Miss:} it \textbf{fails to cross} this beat.
\end{itemize}

\paragraph{Non-summoned Outsider.}
If it has no Leash, create a \textbf{temporary Exit Tally = Cap + 2} the first time it is affected; treat Leash gains as ticks on that tally; when filled, it departs after a brief "acts to nature" beat.

\paragraph{Scope.}
Ward shape, duration, materials, and any advantage/disadvantage are defined by the source Talent/Ability or Spell. \textbf{DV for crossing is always the target's Cap.} \\
\textbf{Wards can affect anything,} but the ability must \textbf{explicitly state} what they affect; the unified DV/Leash math in this section applies only when the target is an \textbf{Outsider}.

\paragraph{\texttt{[UNWARD]} to remove.}
Only an ability with \texttt{[UNWARD]} (or an explicit counter named in the Ward's text) can dismiss a Ward.

\subsection{\texttt{[BANISH]}}\index{Magic!Tags!BANISH}
Drives a visible \textbf{Outsider} toward departure.
\begin{itemize}
  \item Target a visible Outsider in range; roll the pool the source lists.
  \item \textbf{DV = target's Cap.}
  \item \textbf{On a Hit:} add \textbf{+DV segments} to its \textbf{Leash} (or Exit Tally).
  \item \textbf{On a Partial:} add \textbf{+1 segment}.
  \item \textbf{On a Miss:} no effect.
\end{itemize}
If this fills the Leash/Tally, resolve a brief "acts to nature" beat, then the Outsider departs.

\subsection{\texttt{[UNWARD]}}\index{Magic!Tags!UNWARD}
Unmakes or suppresses a \texttt{[WARD]} created by a Talent/Ability or as the result of a Spell.
\begin{itemize}
  \item \textbf{DV:} by fiction (materials, sanctity, prep, locus, opposition).
  \item \textbf{On a Hit:} Ward dismissed/suppressed per text.
  \item \textbf{On a Partial:} weakened or suppressed briefly (about one beat), per GM call.
  \item \textbf{On a Miss:} no effect.
\end{itemize}

\section{One-Glance Summary}\index{Magic!Summary}

\begin{itemize}
  \item \textbf{Outsider} = not native (summoned, demons, extraplanar, spirits, some celestials).
  \item \textbf{Unified DV/impact for Outsiders:} \texttt{[WARD]} crossing and \texttt{[BANISH]} use \textbf{DV = Cap}; \textbf{Hit} $\rightarrow$ \textbf{+DV} Leash segments; \textbf{Partial} $\rightarrow$ \textbf{+1}.
  \item \textbf{Leash} lives on summoned spirits (\textbf{Cap + 2} segments). Non-summoned Outsiders get a temporary \textbf{Exit Tally = Cap + 2} when first affected.
  \item \textbf{Boon Finesse is core:} once per round, spend 1 Boon to clear 1 Leash tick.
  \item \textbf{Gates:} Tags apply only when printed on a \textbf{Talent}, \textbf{Ability}, or as the \textbf{result of a Spell} and cannot be applied with just an action.
  \item \textbf{Summons depart at downtime} unless stated otherwise.
\end{itemize}

\section{The Caster's Burden}\index{Magic!Caster's Burden}

Magicians are defined not by what they can do, but by what they are willing to risk. A cautious spellcaster describes carefully, invests in detailed actions, and may survive long. A reckless one courts power at great personal and narrative cost. Both choices shape the story.

\section{Backlash Severity Table}\index{Magic!Backlash Severity}

\begin{description}
\item[1--2 CP] Minor nuisance or tell; short-lived cost, noise, or reveal.\index{Magic!Backlash!Minor}
\item[3--4 CP] Noticeable setback: a real hazard, condition, or new pressure clock.\index{Magic!Backlash!Moderate}
\item[5+ CP] Major turn: scene shifts, a new foe/clock enters, or severe condition.\index{Magic!Backlash!Major}
\end{description}

\section{Global Guardrails}\index{Magic!Global Guardrails}

\begin{itemize}
\item Duration defaults: buffs 3 beats; areas 1 beat. Sustaining costs 1 Fatigue/beat.\index{Magic!Duration}
\item Stacking: same-Art buffs do not stack; take the best one.\index{Magic!Stacking}
\item Diamond line: spells cannot create Diamonds; only position/effect/rails/clock movement.\index{Magic!Diamond Line}
\item Over-Stack: any standing spell advantage counts toward Over-Stack when the party enters a scene with multiple structural edges.\index{Magic!Over-Stack}
\end{itemize}

\section{Deck-Based Spell Seed Generator}\index{Magic!Spell Generation}

When improvising, draw 2--3 cards. Use the suit to pick an Art, rank to set DV and a scope tweak. Face cards add a quirk; Aces add a price.

\subsection{Suit Arts}\index{Magic!Spell Generation!Suits}
\begin{description}
\item[\textcolor{red}{$\heartsuit$}] Vitalism / Hydromancy: life, vigor, calm, flow, restoration\index{Magic!Spell Generation!Hearts}
\item[\textcolor{black}{$\clubsuit$}] Geomancy / Discipline: stone, structure, weight, binding\index{Magic!Spell Generation!Clubs}
\item[\textcolor{red}{$\diamondsuit$}] Pyromancy / Transformation: heat, light, change, urgency\index{Magic!Spell Generation!Diamonds}
\item[\textcolor{black}{$\spadesuit$}] Umbramancy / Veil: shadow, silence, misdirection, fear\index{Magic!Spell Generation!Spades}
\end{description}

\subsection{Rank $\rightarrow$ DV \& Scope}\index{Magic!Spell Generation!Rank}
\begin{itemize}
\item 2--4: DV 1 (self or tiny area; 1 target; 1 beat sense/edge)\index{Magic!Spell Generation!DV 1}
\item 5--9: DV 2 (small zone $\sim$10 ft; team-sized buff; 1 rail tick)\index{Magic!Spell Generation!DV 2}
\item 10, J, Q, K: DV 3 (loud, group-facing, or multi-rail influence)\index{Magic!Spell Generation!DV 3}
\item Ace: DV 3 + price (component, Fatigue 1, Harm 1, or an Obvious flare)\index{Magic!Spell Generation!Ace}
\end{itemize}

\subsection{Face-card Quirks}\index{Magic!Spell Generation!Face Cards}
\begin{description}
\item[J] Brief mobility or reach boost (leap, glide, slip)\index{Magic!Spell Generation!Jack}
\item[Q] Social resonance (hush, awe, dread)\index{Magic!Spell Generation!Queen}
\item[K] Durability/impact bump (+1 effect on one hard task)\index{Magic!Spell Generation!King}
\end{description}

\subsection{Backlash Prompts by Suit}\index{Magic!Spell Generation!Backlash}
\begin{description}
\item[\textcolor{red}{$\heartsuit$} (Vitalism/Water)] overgrowth, exhaustion, stagnation, sympathetic drain\index{Magic!Spell Generation!Hearts Backlash}
\item[\textcolor{black}{$\clubsuit$} (Stone/Discipline)] rigidity, slow, echoing calls, guardians notice\index{Magic!Spell Generation!Clubs Backlash}
\item[\textcolor{red}{$\diamondsuit$} (Fire/Change)] flare, scorch, noise, uncontrolled spread\index{Magic!Spell Generation!Diamonds Backlash}
\item[\textcolor{black}{$\spadesuit$} (Shadow/Veil)] whispers, self-blindness, cold spots, entities take interest\index{Magic!Spell Generation!Spades Backlash}
\end{description}

\subsection{Generated Example}\index{Magic!Spell Generation!Example}
Drawn: $\diamondsuit$ 10, $\spadesuit$ J

Name: Ember-Stride (Pyromancy). DV: 3.

Effect: For 1 beat, you dash through a chokepoint with start Controlled and +1 effect on clears; leaves a hot shimmer that Hazard --1 as pursuers hesitate. Obvious.

Backlash: 1--2 CP singe/heat-haze reveals route; 3--4 CP smolder opens a Fire(4) clock; 5+ CP flare alarms watchers (start Crowd +1 and Hunt +1).

\section{Play Examples}\index{Magic!Examples}

\subsection{Example 1: Combat Casting}\index{Magic!Examples!Combat}
Situation: Lyra the Pyromancer faces bandits in a narrow alley.
Channel Phase: Lyra rolls Wits + Arcana (5 dice): 8, 6, 4, 1, 1 = 3 successes, 2 CP
GM spends 1 CP: A nearby barrel starts to smolder (minor environmental hazard)
Weave Phase: Next round, Lyra Weaves Cinder-Fist with Detailed description (re-rolls 1s):
Wits + Pyromancy (5 dice): 9, 7, 5, 3, 6 = 4 successes
Effect: Her flaming fist grants +1 effect on melee attacks for 3 beats
GM spends remaining 1 CP: The smoldering barrel bursts into flame, creating a small Fire(4) clock as a hazard

\subsection{Example 2: Utility Casting}\index{Magic!Examples!Utility}
Situation: Theron the Geomancer needs to sense what lies beneath a suspicious floor.
Channel Phase: Theron rolls Wits + Arcana (4 dice): 10, 6, 1, 1 = 2 successes, 2 CP
GM spends 2 CP: The stone whispers of ancient pain, imposing a brief distraction (–1 die on next action)
Weave Phase: Theron Weaves Stone-Sense with Intricate description (re-rolls all 1s and adds flourish): Wits + Geomancy (4 dice): 7, 5, 4, 3 = 3 successes
Effect: He senses a hidden chamber beneath, granting Controlled position for accessing it

\subsection{Example 3: High-Stakes Ritual}\index{Magic!Examples!Ritual}
Situation: The party needs to heal a critically wounded ally during a siege.
Channel Phase: Elena rolls Wits + Arcana (6 dice): 9, 8, 6, 4, 1, 1 = 4 successes, 2 CP
GM spends 1 CP: The life force drawn causes nearby plants to wither
Weave Phase: Elena Weaves Mend Flesh with Detailed description: Wits + Vitalism (6 dice): 10, 7, 5, 3, 2, 1 = 4 successes, 1 CP
Effect: She steps down the ally's Severe injury to Moderate
GM spends remaining CP: The life drawn from surroundings creates an unsettling stillness—small creatures flee the area, and an ominous quiet falls over the battlefield
\end{chapter}
