\chapter{Appendices}

\section{Quick Reference Sheets}

\subsection{Core Mechanic: The Art of Consequence}

\begin{enumerate}
\item \textbf{Approach}: Player states intent and method (Attribute + Skill).
\item \textbf{Execution}: Roll dice pool of \dice{10}s. Each 6+ is a success; each 1 is a Complication Point.
\item \textbf{Outcome}: Basic Roll as-is; Detailed re-roll one 1; Intricate re-roll all 1s and add one positive flourish if successful.
\end{enumerate}

\subsection{Attributes and Skills}

\begin{description}
\item[Body] Strength, endurance, physical force.
\item[Wits] Perception, cleverness, reflexes.
\item[Spirit] Willpower, intuition, resilience.
\item[Presence] Charm, command, social force.
\end{description}

\subsection{Skill Ratings (0--5)}

\begin{description}
\item[0 Untrained] Rely on raw Attribute.
\item[1 Familiar] Basic competence.
\item[2 Skilled] Reliable training.
\item[3 Expert] Professional mastery.
\item[4 Master] Renowned in your field.
\item[5 Legendary] Near-mythic talent.
\end{description}

\subsection{XP Costs}

\begin{description}
\item[Attributes] New rating $\times$ 3.
\item[Skills] New level $\times$ 2.
\item[Followers (On-Screen)] Cap$^2$ XP.
\item[Off-Screen Assets] Minor 4 XP, Standard 8 XP, Major 12 XP.
\end{description}

\section{Deck of Consequences}

Draw from the deck (or roll a d52 equivalence) when Complication Points are spent. Suits = type of complication. Rank = severity.

\subsection{Suit Complication Domains}

\begin{description}
\item[\textcolor{red}{$\heartsuit$}] Emotional / Social fallout (fear, anger, betrayal)
\item[\textcolor{red}{$\diamondsuit$}] Resource / Wealth loss (gear breaks, expenses rise)
\item[\textcolor{black}{$\clubsuit$}] Physical harm / Obstacles (injuries, blockades, fatigue)
\item[\textcolor{black}{$\spadesuit$}] Mystical / Narrative twists (omens, curses, chance)
\end{description}

\subsection{Ranks (1--10, J--K--A)}

\begin{description}
\item[2--5] Minor setback, scene continues smoothly.
\item[6--9] Moderate complication; new obstacle or clock starts.
\item[10--King] Severe twist; alters stakes of the scene.
\item[Ace] Catastrophic turn; reshapes narrative or mission goal.
\end{description}

\section{Magic Casting Loop}

\begin{enumerate}
\item \textbf{Channel}: Wits + Arcana roll to gather Potential.
\item \textbf{Weave}: Wits + Art roll to shape spell.
\item \textbf{Backlash}: Complication Points spent through Deck of Consequences, themed to the opposing element.
\end{enumerate}

\section{Player Archetypes}

\begin{description}
\item[Solo] Invests in Attributes + Skills. Strong spotlight.
\item[Mixed] Balances self with one follower or off-screen assets.
\item[Mastermind] Multiple followers + networks, but more narrative liabilities.
\end{description}

\section{Narrative Time}

\begin{description}
\item[A Moment] A heartbeat; single action.
\item[Some Time] A few minutes; quick exchanges.
\item[Significant Time] An hour or more; downtime actions.
\item[Days] Extended travel or projects.
\end{description}

\section{GM Guidance at a Glance}

\begin{itemize}
\item Spend Complication Points to add story problems, not punish.
\item Always tie consequences back to thematic domains (\textcolor{red}{$\heartsuit$}, \textcolor{red}{$\diamondsuit$}, \textcolor{black}{$\clubsuit$}, \textcolor{black}{$\spadesuit$}).
\item Encourage Intricate actions: reward description with rerolls and narrative control.
\item Let Off-Screen assets resolve downtime problems but keep adventures on the table.
\end{itemize}

\section{Design Guardrails (So It Feels Fair)}

\begin{itemize}
\item Telegraph clocks. Name them aloud or show them in fiction.
\item Spend CP transparently. "I'm using 2 CP to tick Patrol Sweep."
\item Always offer a way out. Boon, asset burn, or devil's bargain.
\item Let Fatigue bite. It's your attrition dial; don't be shy.
\end{itemize}

\section{Clock Starters (d6 Prompts)}

\begin{enumerate}
\item Peril (6): "They triangulate your hideout." Triggers: noise, bribed neighbor, CP spend.
\item Supply (4): "The cold eats your prep." Triggers: storm, lost mule, CP spend.
\item Doom (8): "Blood moon eats the sky." Triggers: daily tick, ritual step, CP spend.
\item Debt (6): "Collector circles." Triggers: downtime passes, flash spend, CP spend.
\item Injury (4): "Cracking ice underfoot." Triggers: sprint, fall, CP spend.
\item Hunt (6): "Rival fixer stalks your routes." Triggers: contact flips, trace, CP spend.
\end{enumerate}

\section{Sample Play (Clocks + CP + Fatigue)}

\subsection{Scene Frame}
Smuggling a witness over the palace wall at night. Active clocks: Patrol Sweep (6) at 3/6, Supply (4) at 1/4.

\subsection{Beats}

\textbf{1) Scout the south wall} Wits+Stealth (5\dice{10}): 10, 8, 6, 3, 1 $\Rightarrow$ 3 hits, 1 CP. GM spends 1 CP: "You succeed, but your old rope is Compromised." Patrol Sweep stays at 3/6.

\textbf{2) Climb with the witness} Body+Athletics (4\dice{10}): 7, 6, 5, 1 $\Rightarrow$ 2 hits, 1 CP. GM spends 1 CP to tick Patrol Sweep $\Rightarrow$ 4/6 and adds pressure: "Lanterns brighten; bootsteps quicken along the inner walk."

\textbf{3) Last push, running low} GM fills Supply to 2/4 due to Significant Time climbing and cold. Face tries to soothe the witness (Presence+Sway, 5\dice{10}): 9, 8, 2, 1, 1 $\Rightarrow$ 2 hits, 2 CP. GM spends 2 CP: +1 Fatigue to the Face (shivering, breathless) and tick Patrol Sweep to 5/6.

\textbf{4) Extraction or collapse} They have a choice: burn a Boon to activate an off-screen contact's garden gate, or risk one more roll with Fatigue penalties. If they roll and give the GM 1 CP, the GM can fill Patrol Sweep to 6/6 $\Rightarrow$ "Cordon slams shut; horns blare. Capture is on the table."

\section{Nomenclature (Quick Canon)}

\subsection{Regions and Peoples}
\begin{itemize}
\item Kahfagia --- western convoys lantern law.
\item Ecktoria --- central peninsula; coinhouses forums.
\item Acasia --- broken marches; Silkstrand the sole metropolis and Amaranthine hub.
\item Vhasia --- fractured sun; Marcott at the Belworth mouth.
\item Viterra --- last kingdom; Fairport at the Belworth mouth; Tarlington on the falls.
\item Ubral --- stone between spears; passes to the Dolmis NW.
\item Thepyrgos --- stairs, synod, and the Astroegro Straits.
\item Aeler --- crowns under-vaults across the Aelerians.
\item Mistlands --- bells, salt, and breath; Payden's Port.
\item Valewood --- empire under leaves; thin shore.
\item Aelinnel --- stone, bough, and bright tide-gates.
\item Aelaerem --- hearth hollow across rolling downs.
\item Linn --- skerries storm-oaths; the Yloka road.
\item Ykrul --- wolf standards winter camps.
\item Vilikari --- laurels longhouses.
\end{itemize}

\subsection{Geography}
\begin{itemize}
\item Amaranthine Sea --- western sea.
\item Dolmis Sea --- inner sea to the east.
\item Aberderrin Sea --- northeastern basin.
\item Astroegro Straits --- hinge between seas.
\end{itemize}

