\chapter{Lore-Heavy Backgrounds, Skills, \& Talents}

\section{Skills}\index{Skills!Lore-Heavy}

\subsection{Skill: Utaran Law \& Precedent (Intricate)}\index{Skills!Utaran Law}
\textbf{Scope:} Imperial codes, edicts, and forum custom from Ecktoria to the marches.
\textbf{Approach:} Wits or Presence + Utaran Law.
\textbf{Effect:} On success in a legal/bureaucratic scene, you may downgrade one \emph{legal} \SuitClub{} in this leg to \emph{Bureaucratic Delay} (buys time without removing scrutiny).
\textbf{Crit:} Mint a temporary \SuitDiamond{} \emph{Stay of Writ} (Debt Flip: a clerk demands a future favor).
\textbf{Complications (CP spend):} audit flags, paperwork trail, or offended minor official.
\textbf{Deck hook:} In Vhasia/Viterra ports, converts \emph{quarantine/levy} Clubs to \emph{inspection} (softer bite).

\subsection{Skill: Stone-Tongue (Aeler Affinity)}\index{Skills!Stone-Tongue}
\textbf{Scope:} Read the memory of stone, unworked rock to vault masonry.
\textbf{Approach:} Spirit + Stone-Tongue (Intricate if you describe the touching rite).
\textbf{Effect:} Ask the GM 1--3 questions about \emph{who carved/used/strained} this stone. Each extra question risks 1 CP of \emph{haunt} (fatigue, intrusive memory).
\textbf{Travel rider:} Once per Aeler leg, you may peek at the next leg's \SuitSpade{} or \SuitClub{}.
\textbf{Complications:} Disorientation; GM may add a subtle echo/omen to later scenes.

\subsection{Skill: Monsoon-Reckoning (Southern/Eastern Belts)}\index{Skills!Monsoon-Reckoning}
\textbf{Scope:} Seasonal winds and their social/economic knock-on effects (Akilan/Ostrilan routes).
\textbf{Approach:} Wits + Monsoon-Reckoning (Intricate if you read skies, currents, ledgers).
\textbf{Effect:} Once per sea/river leg, shift a \emph{weather} \SuitClub{} down one rank (A→10, 10→9, etc.) \emph{or} convert it to a \emph{Forecast} boon (DV $-1$ to the next two navigation/planning rolls this leg).
\textbf{Complications:} If you miss by 2+, GM banks +1 CP and introduces a \emph{false window} later.

\subsection{Skill: Heraldry \& Lineage (Vhasia/Vilikari)}\index{Skills!Heraldry}
\textbf{Scope:} Houses, devices, marriages, feuds.
\textbf{Approach:} Presence + Heraldry.
\textbf{Effect:} Gain \emph{Controlled} position vs titled NPCs; ask 1 specific House-Truth on success.
\textbf{Crit:} One-use \emph{Letters of Lineage} pass (acts as a local Warrant \SuitDiamond{}).
\textbf{Complications:} Name a feud; someone you slighted takes notice.

\subsection{Skill: Geasa \& Oath-Weaving (Theona/Elven)}\index{Skills!Geasa}
\textbf{Scope:} Binding promises, taboos, and loopholes.
\textbf{Approach:} Spirit + Geasa (Intricate if you state exact terms, witness, and token).
\textbf{Effect:} Attach a 4-segment \emph{geas} clock to a willing target; define trigger \& consequence (GM must honor).
\textbf{Fail/Complications:} GM gains +2 CP as \emph{oath backlash}, or you catch a minor bind until sunset.
\textbf{Deck hook:} Theona/Valewood \emph{9s} can satisfy or complicate a geas condition.

\section{Talents}\index{Talents!Lore-Heavy}

\subsection{Talent (Prestige): Echo-Walker's Step}\index{Talents!Echo-Walker's Step}
\textbf{Req:} Visit a site of deep memory (ruined Lethai-ar city, great betrayal, etc.).
\textbf{Use:} 1/arc, enter an echo of a past event at your location. Observe perfectly; you may handle ephemeral objects.
\textbf{Cost/Complications:} GM immediately banks +2 CP; a figure in the echo may \emph{notice} you (complication tag).
\textbf{Synergy:} Treat the next \emph{mystery} \SuitClub{} in this arc as already \emph{Known} for DV purposes (DV $-1$).

\subsection{Talent (Mid): Margravine's Warrant (Acasia)}\index{Talents!Margravine's Warrant}
\textbf{Req:} Complete a job for a petty ruler.
\textbf{Use:} Carry a portable \SuitDiamond{} \emph{Warrant} of rank (7--K). Spend to bypass a minor toll, commandeer limited aid, or sway a magistrate in Acasia.
\textbf{Exposure:} Each use ticks Patron Exposure (3 boxes). At Burned, the warrant is revoked and you're wanted.

\subsection{Talent (Mid): Pilot's Mirror Glint (Kahfagia)}\index{Talents!Pilot's Mirror Glint}
\textbf{Req:} Survive a hazardous channel using lantern-law.
\textbf{Use:} 1/session predict an imminent lantern-law shift; bank +1 free success on a pilotage/nav action tied to that shift.
\textbf{Miss:} GM gains +2 CP; your timing is off with the convoy for this leg.

\subsection{Talent (Prestige): Speak with the Pale Shepherd (Aelaerem)}\index{Talents!Pale Shepherd}
\textbf{Req:} Hearthbound tie (homestead/family) and resolve a hearth threat.
\textbf{Rite:} Ask one question about cycles, loss, or protection; GM must answer plainly (in theme).
\textbf{Cost:} Significant personal offering; you gain the \emph{Melancholy} tag until next dawn (narrative weight).

\subsection{Talent (Prestige): Forge-Seal of Khaz-Vurim (Aeler)}\index{Talents!Forge-Seal}
\textbf{Req:} Craft a masterwork in a major Aeler forge.
\textbf{Effect:} Auto-\emph{Controlled} with dwarven craft-guilds; 1/session upgrade a relevant action to \emph{Intricate} when you demonstrate method on-screen.
\textbf{Downside:} Any deception/shoddy work about craft revokes the Seal (hard loss).

\section{Backgrounds}\index{Backgrounds!Lore-Heavy}
Each background below packages a cultural role with a signature \emph{Talent}, a per-session \emph{Perk choice}, keyed \emph{Scene Keys}, and CP-driven \emph{Hooks}. They integrate with the Description Ladder (sell the fiction to reach \emph{Intricate}), travel clocks (\ClockSizes), and the suit logic: \SuitSpade{}=place, \SuitHeart{}=actor, \SuitClub{}=pressure, \SuitDiamond{}=leverage.

\subsection{Sepulcher Adept (Aeler)}\index{Backgrounds!Sepulcher Adept}
\textbf{Requires Asset:} Aeler Crypt Access \emph{or} Oath-Pennant Keeper.
\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Whisper of the Unquiet Dead (6 XP)}]
Once/session, commune with spirits bound to a named crypt/barrow/under-way. Ask \textbf{one} direct question; answer is truthful but limited to their death and the stone nearby. If you press further, GM may bank +1 CP as \emph{haunt}.
\end{tcolorbox}
\textbf{Perk (choose 1, once/session):}
\begin{itemize}
  \item Mitigate the first CP from a \emph{spiritual/ancestral} \SuitClub{} this scene.
  \item Silence your group's movement for one beat in a sealed stone space.
  \item Instantly detect necromancy or grave-robbing on a corpse or bier.
\end{itemize}
\textbf{Scene Keys:} Royal Crypt; Under-Vault Archive; Disputed Burial Site.
\textbf{Hooks (GM on CP spend in keyed scenes):}
\begin{itemize}
  \item Ancestral spirit misidentifies you; demands recompense or service.
  \item Senior Adept deems your rite overreaching; censure or audit follows.
  \item You learn a truth damning to a living house; secrecy becomes a burden.
\end{itemize}
\textbf{Invite:} "Consult the stones before acting."

\subsection{Lantern-Law Advocate (Kahfagia)}\index{Backgrounds!Lantern-Law Advocate}
\textbf{Requires Asset:} Kahfagian Harbor Charter \emph{or} Pilot's Guild Membership.
\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Jurisdiction on the Tide (5 XP)}]
Once/session in a maritime dispute, shift the ruling one step in your favor (guilty→fine, fine→warning) by citing precedent. If contested by counsel, roll Wits/Presence + Utaran Law; on a miss the GM banks +1 CP and flags a future audit.
\end{tcolorbox}
\textbf{Perk (choose 1, once/session):}
\begin{itemize}
  \item Predict the next beacon shift for a lane; gain one-beat advantage on pilotage.
  \item A minor port official (Cap 2) backs your claim on a procedural technicality.
  \item Halve a customs levy/toll through flawless paperwork.
\end{itemize}
\textbf{Scene Keys:} Admiralty Court; Fog-bound Harbor; Pilot's Guildhall.
\textbf{Hooks (GM on CP spend in keyed scenes):}
\begin{itemize}
  \item Rival cites deeper bylaw; your edge is nullified this scene.
  \item Your precedent is weaponized by a hostile faction immediately.
  \item Lanterns malfunction, proving you right but creating a physical Hazard.
\end{itemize}
\textbf{Invite:} "Quote chapter and verse."

\subsection{Curse-Breaker of Silkstrand (Acasia)}\index{Backgrounds!Curse-Breaker}
\textbf{Requires Asset:} Silkstrand Safehouse \emph{or} Archivolt Research Rights.
\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Follow the Unspooled Thread (7 XP)}]
Once/session when the Curse manifests (looping road, missing ninth, etc.), roll Wits + Lore. On success, reveal one temporary loophole/safe path. On crit, also grant DV $-1$ to the next two allied actions exploiting it. On failure, GM banks +2 CP as \emph{curse backlash}.
\end{tcolorbox}
\textbf{Perk (choose 1, once/session):}
\begin{itemize}
  \item Anchor a location against the Curse for one scene (normal navigation).
  \item Identify a person/object currently anchoring a knot of the Curse.
  \item +1 effect to protect others from indirect Curse fallout.
\end{itemize}
\textbf{Scene Keys:} Cursed Crossroads; Affected Estate; Archivolt Library.
\textbf{Hooks (GM on CP spend in keyed scenes):}
\begin{itemize}
  \item The Curse twists a future success into a failure.
  \item A cult of the Curse marks you for meddling.
  \item Break one knot; two tighten elsewhere, opening a harder leg.
\end{itemize}
\textbf{Invite:} "Trace the fault line."

\subsection{Voice of the Moot (Linn / Theona)}\index{Backgrounds!Voice of the Moot}
\textbf{Requires Asset:} Thing-holm Speaking Rights \emph{or} Foster-Bond with a Jarl/Chieftain.
\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Word-Bond (6 XP)}]
Once/session, oversee a formal oath between parties. If broken, impose a narrative penalty appropriate to culture (ostracism, ill-sea, loss of face) until amends are made. Roll Presence + Geasa/Heraldry as needed to mediate.
\end{tcolorbox}
\textbf{Perk (choose 1, once/session):}
\begin{itemize}
  \item Calm a heated crowd with proverb/history; set Position to Controlled.
  \item Call a minor boon from a party bound by your past oath.
  \item Instantly know if a speaker is under geas/compulsion during parley.
\end{itemize}
\textbf{Scene Keys:} Thing Gathering; Wedding Feast; Oath-Swearing.
\textbf{Hooks (GM on CP spend in keyed scenes):}
\begin{itemize}
  \item A listener exploits an unintended loophole.
  \item You must judge a case where both sides are right; choose who you lose.
  \item An old oath conflicts with the new, nullifying both until reconciled.
\end{itemize}
\textbf{Invite:} "Seek the consensus."

\subsection{Legate of the Silent Gate (Mistlands)}\index{Backgrounds!Legate of the Silent Gate}
\textbf{Requires Asset:} Protectorate Commission \emph{or} Bell-Warden Authority.
\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Breath-Test (8 XP)}]
Once/session, perform the Aeler rite to diagnose a breach: name the incursion type (wraiths, intruders, miasma) and its approximate strength. On a risky read, the GM may bank +1 CP and demand full protocol (time/favor).
\end{tcolorbox}
\textbf{Perk (choose 1, once/session):}
\begin{itemize}
  \item Command a Cap 3 Protectorate squad to secure a point.
  \item Sanctify a small area with ward-salt and bells; safe vs wraiths for one scene.
  \item Decode bell-logs to know quiet/hot watches exactly.
\end{itemize}
\textbf{Scene Keys:} Weeping Gate; Bell-Line Levee; Breach Site.
\textbf{Hooks (GM on CP spend in keyed scenes):}
\begin{itemize}
  \item Local Mistlander challenges Aeler authority.
  \item New horror exceeds protocol; write a rule on the fly.
  \item Procedure demands sacrifice (time, person, or asset) for the greater safety.
\end{itemize}
\textbf{Invite:} "Trust the bells."

\section{Prestige Backgrounds (Epic)}\index{Backgrounds!Prestige}

Each Prestige Background is a destiny-tier role earned in play. Its signature Talent is \textbf{once per campaign}. Using it always creates or advances a Front and triggers the listed Epic Hook. Suits: \SuitSpade=place, \SuitHeart=actor, \SuitClub=pressure, \SuitDiamond=leverage.

\subsection{The Cursed Child of Silkstrand}\index{Backgrounds!Cursed Child}
\textbf{Requires:} Personally mitigate or trigger a major Acasian Curse event; deep tie to Silkstrand.

\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Laughter That Ends Sieges (20 XP, 1/campaign)}]
In a moment of absolute desperation, focus the Curse through yourself. \textbf{End one large-scale conflict} in sight (siege, battle, storm) as reality \emph{unravels} it: engines fail harmlessly, zeal cools, spells fizzle into nonsense. \textbf{No roll.}
\end{tcolorbox}

\textbf{Epic Hook — The Curse Hungers:} The Curse binds to you. At the start of every session, the GM introduces a minor, personal echo (an omission, loop, or wrong reflection) affecting you or allies. \textbf{GM:} bank +2 CP and start/advance a 6-segment \emph{Silkstrand Echo} Front.

\textbf{Deck hooks:} Treat any Valewood/Theona \emph{9} drawn in your scenes as \emph{yours} (the omission centers on you). Spending a local \SuitDiamond{} (e.g., Way-cord) can postpone, not erase, an echo.

\subsection{Stone-Speaker of Khaz-Vurim}\index{Backgrounds!Stone-Speaker}
\textbf{Requires:} Broker a monumental peace between Aeler holds or with a major power; dwarven culture or unprecedented trust.

\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Move the Mountain's Will (20 XP, 1/campaign)}]
Call the deep will of the Aelerians to \textbf{reroute a geological or political reality}: open a pass weeks early, flood a rival shaft, or cause a hold to ignore a king's edict. \textbf{No roll.}
\end{tcolorbox}

\textbf{Epic Hook — The Mountain's Price:} You swear a \emph{Geas of Stone}. The GM assigns a monumental, slow task against your interests. Until completed, you lose dwarven hospitality and stone-magic benefit (treat social Position as \emph{Desperate} with Aeler guilds).

\textbf{Deck hooks:} When used, flip any Aeler \SuitDiamond{} played this arc immediately (its Debt Flip is collected now). Start/advance an 8-segment \emph{Geas of Stone} Front.

\subsection{The Last Dauphin's Shadow}\index{Backgrounds!Dauphin's Shadow}
\textbf{Requires:} Publicly champion, protect, or expose the Sun Court claimant; irrevocably entangled in the cause.

\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — The Sunburst Warrant (18 XP, 1/campaign)}]
Issue commands \textbf{obeyed as if from the lost Sun-King}: mobilize hidden cells, muster a county levy, or claim noble sanctuary. Choose one concrete objective; it happens, unless physically impossible. \textbf{No roll.}
\end{tcolorbox}

\textbf{Epic Hook — The Pretender's Toll:} Each use \textbf{consumes hope}. GM selects a loyalist asset to be lost, disillusioned, or exposed (spy executed, relic destroyed, populace turns). Start/advance a 6-segment \emph{Candle of the Sun} Front; when it fills, the movement fractures.

\textbf{Deck hooks:} In Vhasia, treat face-card \SuitHeart{}s as predisposed (Controlled Position) but \SuitClub{} \emph{legal/economic} hits escalate faster (GM may spend 1 CP as 2 CP against you once/scene).

\subsection{Mist-Seer of the Weeping Gate}\index{Backgrounds!Mist-Seer}
\textbf{Requires:} Endure the Direwood and return changed; face a horror beyond common wraiths.

\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — Bargain with the Deep Mist (20 XP, 1/campaign)}]
Open parley with the Mist's sentience. Ask \textbf{one true question} \emph{or} command it to \textbf{withhold influence} from a named place/person for a year and a day. \textbf{No roll.}
\end{tcolorbox}

\textbf{Epic Hook — The Wood Remembers:} Pay a \emph{Bargain of Breath}: surrender a fundamental memory/name/sense/skill (permanent, narrative). GM banks +2 CP and starts a 6-segment \emph{Lost Thing} Front keyed to what was given up; scenes that touch it impose DV +1 on you until closure.

\textbf{Deck hooks:} A Mistlands \SuitDiamond{} (\emph{Ward-salt} or \emph{Lantern Writ}) can soften—not negate—the next echo of the loss.

\subsection{Keeper of the Foedus Stone}\index{Backgrounds!Keeper of the Foedus Stone}
\textbf{Requires:} Primary mediator preventing a Ykrul/Vilikari (or equivalent) war from breaking.

\begin{tcolorbox}[enhanced,sharp corners,boxrule=.6pt,title={Talent — The Law That Binds Spears (18 XP, 1/campaign)}]
At the Foedus Stone (or an ancient treaty site), \textbf{impose an absolute truce} between two named factions. For its span, any violent act across the line triggers immediate supernatural calamity on the offender. \textbf{No roll.}
\end{tcolorbox}

\textbf{Epic Hook — The Balance of Grudges:} The hate is compressed, not erased. Begin an 8-segment \emph{Coiled Blood} Front that passively ticks each session (or on Club escalations). When it blows, the rematch is worse—your name is on the blame.

\textbf{Deck hooks:} During the truce, \SuitDiamond{}s like \emph{Foedus Seal}, \emph{Paiza}, or \emph{Standard Protection} gain +1 practical scope (broader safe-conduct), but any \emph{hostage} \SuitClub{} automatically targets your allies first on CP spend.
\end{chapter}
