\chapter{Appendices}

\section{Quick Reference Sheets}\index{Appendices!Quick Reference}

\subsection{Core Mechanic: The Art of Consequence}\index{Appendices!Core Mechanic}

\begin{enumerate}
\item \textbf{Approach}: Player states intent and method (Attribute + Skill).
\item \textbf{Execution}: Roll dice pool of \dice{10}s. Each 6+ is a success; each 1 is a Complication Point.
\item \textbf{Outcome}: Basic Roll as-is; Detailed re-roll one 1; Intricate re-roll all 1s and add one positive flourish if successful.
\end{enumerate}

\subsection{Attributes and Skills}\index{Appendices!Attributes and Skills}

\begin{description}
\item[Body] Strength, endurance, physical force.\index{Attributes!Body}
\item[Wits] Perception, cleverness, reflexes.\index{Attributes!Wits}
\item[Spirit] Willpower, intuition, resilience.\index{Attributes!Spirit}
\item[Presence] Charm, command, social force.\index{Attributes!Presence}
\end{description}

\subsection{Skill Ratings (0--5)}\index{Appendices!Skill Ratings}

\begin{description}
\item[0 Untrained] Rely on raw Attribute.\index{Skills!Untrained}
\item[1 Familiar] Basic competence.\index{Skills!Familiar}
\item[2 Skilled] Reliable training.\index{Skills!Skilled}
\item[3 Expert] Professional mastery.\index{Skills!Expert}
\item[4 Master] Renowned in your field.\index{Skills!Master}
\item[5 Legendary] Near-mythic talent.\index{Skills!Legendary}
\end{description}

\subsection{XP Costs}\index{Appendices!XP Costs}

\begin{description}
\item[Attributes] New rating $\times$ 3.\index{Experience Points!Attributes}
\item[Skills] New level $\times$ 2.\index{Experience Points!Skills}
\item[Followers (On-Screen)] Cap$^2$ XP.\index{Experience Points!Followers}
\item[Off-Screen Assets] Minor 4 XP, Standard 8 XP, Major 12 XP.\index{Experience Points!Assets}
\end{description}

\section{Deck of Consequences}\index{Appendices!Deck of Consequences}

Fate's Edge uses two distinct card tools:

\paragraph{Travel Decks (regional, 52-card).}\index{Decks!Travel}
\emph{Spade}=Place, \emph{Heart}=Actor, \emph{Club}=Pressure, \emph{Diamond}=Leverage. These power journeys and gates.

\paragraph{Deck of Consequences (scene drama).}\index{Decks!Consequences}
\emph{Hearts}=social fallout, \emph{Spades}=harm/escalation, \emph{Clubs}=material cost, \emph{Diamonds}=magical/spiritual disturbance.

\textit{Guidance:} Never mix suit meanings across decks. When a rule references ``Spade/Club/Diamond,'' it means \emph{Travel}. When it says ``Hearts/Spades/Clubs/Diamonds,'' it means \emph{Consequences}.

\subsection{Using the Deck}\index{Appendices!Using the Deck}

After a roll that generates CP, the GM chooses one method for that roll:
\begin{enumerate}
  \item \textbf{Direct Spend}: Translate CP into consequences/rail ticks immediately.
  \item \textbf{Deck Draw}: Draw up to \textbf{min(CP, 3)} cards and synthesize a single twist guided by suit and highest rank.
\end{enumerate}

\subsection{Consequences Deck Structure}\index{Appendices!Consequences Deck}

\begin{description}
\item[Hearts] Emotional / Social fallout (fear, anger, betrayal)\index{Decks!Consequences!Hearts}
\item[Spades] Physical harm / Obstacles (injuries, blockades, fatigue)\index{Decks!Consequences!Spades}
\item[Clubs] Resource / Wealth loss (gear breaks, expenses rise)\index{Decks!Consequences!Clubs}
\item[Diamonds] Mystical / Narrative twists (omens, curses, chance)\index{Decks!Consequences!Diamonds}
\end{description}

\subsection{Ranks (1--10, J--K--A)}\index{Appendices!Ranks}

\begin{description}
\item[Ace--3] Minor inconvenience or flavor complication.\index{Decks!Ranks!Minor}
\item[4--6] Moderate setback with some narrative teeth.\index{Decks!Ranks!Moderate}
\item[7--9] Significant consequence altering the course of action.\index{Decks!Ranks!Significant}
\item[10--King] Major fallout, introducing new problems or lasting scars.\index{Decks!Ranks!Major}
\end{description}

\section{Magic Casting Loop}\index{Appendices!Magic}

\begin{enumerate}
\item \textbf{Channel}: Wits + Arcana roll to gather Potential.\index{Magic!Channel}
\item \textbf{Weave}: Wits + Art roll to shape spell.\index{Magic!Weave}
\item \textbf{Backlash}: Complication Points spent through Deck of Consequences, themed to the opposing element.\index{Magic!Backlash}
\end{enumerate}

\textbf{Mitigation}: Boons do not reduce CP unless a Talent/Asset explicitly says "Mitigate CP."\index{Boons!Mitigation}

\section{Player Archetypes}\index{Appendices!Player Archetypes}

\begin{description}
\item[Solo] Invests in Attributes + Skills. Strong spotlight.\index{Character Build!Solo}
\item[Mixed] Balances self with one follower or off-screen assets.\index{Character Build!Mixed Player}
\item[Mastermind] Multiple followers + networks, but more narrative liabilities.\index{Character Build!Mastermind}
\end{description}

\section{Narrative Time}\index{Appendices!Narrative Time}

\begin{description}
\item[A Moment] A heartbeat; single action.\index{Narrative Time!Moment}
\item[Some Time] A few minutes; quick exchanges.\index{Narrative Time!Some Time}
\item[Significant Time] An hour or more; downtime actions.\index{Narrative Time!Significant Time}
\item[Days] Extended travel or projects.\index{Narrative Time!Days}
\end{description}

\section{GM Guidance at a Glance}\index{Appendices!GM Guidance}

\begin{itemize}
\item Spend Complication Points to add story problems, not punish.\index{Complication Points!GM Spending}
\item Always tie consequences back to thematic domains (Hearts, Spades, Clubs, Diamonds).\index{Decks!Consequences}
\item Encourage Intricate actions: reward description with rerolls and narrative control.\index{Description Ladder}
\item Let Off-Screen assets resolve downtime problems but keep adventures on the table.\index{Assets!Off-Screen}
\end{itemize}

\section{Design Guardrails (So It Feels Fair)}\index{Appendices!Design Guardrails}

\begin{itemize}
\item Telegraph clocks. Name them aloud or show them in fiction.\index{Clocks!Telegraphing}
\item Spend CP transparently. "I'm using 2 CP to tick Patrol Sweep."\index{Complication Points!Transparent Spending}
\item Always offer a way out. Boon, asset burn, or devil's bargain.\index{Complication Points!Devil's Bargain}
\item Let Fatigue bite. It's your attrition dial; don't be shy.\index{Fatigue}
\end{itemize}

\section{Mechanical Constraints}\index{Appendices!Mechanical Constraints}
\begin{itemize}
    \item \textbf{ASSIST MAX}: +3 dice total per roll, regardless of helpers. Exception: The "Exceptional Coordination" Talent allows one follower to provide +4 assist dice.\index{Mechanics!Assist Maximum}
    \item \textbf{BOON MAX}: 5 total, 2→1 XP conversion once/session (max 2 XP via conversion per session).\index{Boons!Maximum}
    \item \textbf{INITIATIVE}: 1 Follower Action per scene party-wide.\index{Combat!Follower Initiative}
    \item \textbf{OVER-STACK}: 2+ structural advantages = start rails +1 OR GM banks +1 CP.\index{Combat!Over-Stack}
    \item \textbf{POSITION}: Controlled | Risky | Desperate (affects success/failure texture).\index{Combat!Position States}
\end{itemize}

\section{Clock Starters (d6 Prompts)}\index{Appendices!Clock Starters}

\begin{enumerate}
\item Peril (6): "They triangulate your hideout." Triggers: noise, bribed neighbor, CP spend.\index{Clocks!Peril}
\item Supply (4): "The cold eats your prep." Triggers: storm, lost mule, CP spend.\index{Clocks!Supply}
\item Doom (8): "Blood moon eats the sky." Triggers: daily tick, ritual step, CP spend.\index{Clocks!Doom}
\item Debt (6): "Collector circles." Triggers: downtime passes, flash spend, CP spend.\index{Clocks!Debt}
\item Injury (4): "Cracking ice underfoot." Triggers: sprint, fall, CP spend.\index{Clocks!Injury}
\item Hunt (6): "Rival fixer stalks your routes." Triggers: contact flips, trace, CP spend.\index{Clocks!Hunt}
\end{enumerate}

\section{Sample Play (Clocks + CP + Fatigue)}\index{Appendices!Sample Play}

\subsection{Scene Frame}\index{Appendices!Sample Play!Scene Frame}
Smuggling a witness over the palace wall at night. Active clocks: Patrol Sweep (6) at 3/6, Supply (4) at 1/4.

\subsection{Beats}\index{Appendices!Sample Play!Beats}

\textbf{1) Scout the south wall} Wits+Stealth (5\dice{10}): 10, 8, 6, 3, 1 $\Rightarrow$ 3 hits, 1 CP. GM spends 1 CP: "You succeed, but your old rope is Compromised." Patrol Sweep stays at 3/6.

\textbf{2) Climb with the witness} Body+Athletics (4\dice{10}): 7, 6, 5, 1 $\Rightarrow$ 2 hits, 1 CP. GM spends 1 CP to tick Patrol Sweep $\Rightarrow$ 4/6 and adds pressure: "Lanterns brighten; bootsteps quicken along the inner walk."

\textbf{3) Last push, running low} GM fills Supply to 2/4 due to Significant Time climbing and cold. Face tries to soothe the witness (Presence+Sway, 5\dice{10}): 9, 8, 2, 1, 1 $\Rightarrow$ 2 hits, 2 CP. GM spends 2 CP: +1 Fatigue to the Face (shivering, breathless) and tick Patrol Sweep to 5/6.

\textbf{4) Extraction or collapse} They have a choice: burn a Boon to activate an off-screen contact's garden gate, or risk one more roll with Fatigue penalties. If they roll and give the GM 1 CP, the GM can fill Patrol Sweep to 6/6 $\Rightarrow$ "Cordon slams shut; horns blare. Capture is on the table."

\section{Nomenclature (Quick Canon)}\index{Appendices!Nomenclature}

\subsection{Regions and Peoples}\index{Appendices!Nomenclature!Regions}
\begin{itemize}
\item Kahfagia --- western convoys lantern law.\index{Kahfagia}
\item Ecktoria --- central peninsula; coinhouses forums.\index{Ecktoria}
\item Acasia --- broken marches; cursed and lawless territory.\index{Acasia}
\item Silkstrand --- port city of trade, secrets, bravos, and intrigue; sole major city in Acasia on the Amaranthine Sea.\index{Silkstrand}
\item Vhasia --- fractured sun; Marcott at the Belworth mouth.\index{Vhasia}
\item Viterra --- last kingdom; Fairport at the Belworth mouth; Tarlington on the falls.\index{Viterra}
\item Ubral --- stone between spears; passes to the Dolmis NW.\index{Ubral}
\item Thepyrgos --- stairs, synod, and the Astroegro Straits.\index{Thepyrgos}
\item Aeler --- crowns under-vaults across the Aelerians.\index{Aeler}
\item Mistlands --- bells, salt, and breath; Payden's Port.\index{Mistlands}
\item Valewood --- empire under leaves; thin shore.\index{Valewood}
\item Aelinnel --- stone, bough, and bright tide-gates.\index{Aelinnel}
\item Aelaerem --- hearth hollow across rolling downs.\index{Aelaerem}
\item Linn --- skerries storm-oaths; the Yloka road.\index{Linn}
\item Ykrul --- wolf standards winter camps.\index{Ykrul}
\item Vilikari --- laurels longhouses.\index{Vilikari}
\item Zakov --- salt serpent corsairs; criminal syndicates and smuggler's haven on the Dolmis coast.\index{Zakov}
\item Black Banners --- condotta free companies; mercenaries and war-band territories in Ykrul lands.\index{Black Banners}
\end{itemize}

\subsection{Geography}\index{Appendices!Nomenclature!Geography}
\begin{itemize}
\item Amaranthine Sea --- western sea.\index{Geography!Amaranthine Sea}
\item Dolmis Sea --- inner sea to the east.\index{Geography!Dolmis Sea}
\item Aberderrin Sea --- northeastern basin.\index{Geography!Aberderrin Sea}
\item Astroegro Straits --- hinge between seas.\index{Geography!Astroegro Straits}
\end{itemize}

\section{Prestige Abilities}\index{Appendices!Prestige Abilities}
\begin{itemize}
    \item \textbf{Echo-Walker's Step} (High Elf, Cost: 20 XP; Req: Wits 5, Arcana 4): 
1/arc, \emph{observe} a perfect echo of a past event at your location (no retconning). 
GM immediately banks +2 CP; scenes touching that memory carry an omen. Grants DV −1 on one action that uses the revealed truth.\index{Prestige Abilities!Echo-Walker's Step}
    \item \textbf{Warglord} (Ykrul, Cost: 18 XP; Req: Body 5, Command 3): 
Once per campaign, unify scattered warbands into a single host for a season. Start a \emph{Logistics} clock and a \emph{Grudge} clock; either one filling fractures the host.\index{Prestige Abilities!Warglord}
    \item \textbf{Spirit-Shield} (Aeler, Cost: 15 XP; Req: Spirit 4, Insight 3): 
1/session, erase up to 3 CP from an ally's \emph{current} roll; you immediately mark Fatigue +1 and the GM banks +1 CP as backlash.\index{Prestige Abilities!Spirit-Shield}
\end{itemize}

\section{Between Sessions Activities Log}\index{Appendices!Session Log}

\subsection{Character Advancement}\index{Appendices!Session Log!Character Advancement}

\subsubsection{XP Allocation}\index{Appendices!Session Log!XP}
\begin{itemize}
    \item \textbf{Total XP Available}: \_\_\_\_\_ XP
    \item \textbf{Attributes Spent}: \_\_\_\_\_ XP (\_\_\_\_\_ days downtime)
    \item \textbf{Skills Spent}: \_\_\_\_\_ XP (\_\_\_\_\_ days downtime)
    \item \textbf{Remaining XP}: \_\_\_\_\_ XP
\end{itemize}

\subsubsection{Attribute Improvements}\index{Appendices!Session Log!Attributes}
\begin{tabular}{|p{3cm}|c|c|c|c|}
\hline
\textbf{Attribute} & \textbf{Old Rating} & \textbf{New Rating} & \textbf{Cost} & \textbf{Downtime} \\
\hline
Might & & & & \\
Finesse & & & & \\
Resolve & & & & \\
Insight & & & & \\
Aura & & & & \\
\hline
\end{tabular}

\subsubsection{Skill Improvements}\index{Appendices!Session Log!Skills}
\begin{tabular}{|p{4cm}|c|c|c|c|}
\hline
\textbf{Skill} & \textbf{Old Level} & \textbf{New Level} & \textbf{Cost} & \textbf{Downtime} \\
\hline
Academia & & & & \\
Arts & & & & \\
Athletics & & & & \\
Bonds & & & & \\
Combat & & & & \\
Craft & & & & \\
Deceive & & & & \\
Empathy & & & & \\
Lore & & & & \\
Move & & & & \\
Notice & & & & \\
Perform & & & & \\
Physique & & & & \\
Pilot & & & & \\
Ranged & & & & \\
Rapport & & & & \\
Resist & & & & \\
Stealth & & & & \\
Tactics & & & & \\
Will & & & & \\
\hline
\end{tabular}

\subsection{Asset Management}\index{Appendices!Session Log!Assets}

\subsubsection{New Assets Acquired}\index{Appendices!Session Log!New Assets}
\begin{tabular}{|p{3cm}|p{2cm}|p{2cm}|p{3cm}|}
\hline
\textbf{Asset Name} & \textbf{Type} & \textbf{Cost} & \textbf{Acquisition Method} \\
\hline
& & & \\
& & & \\
& & & \\
\hline
\end{tabular}

\subsubsection{Existing Assets}\index{Appendices!Session Log!Existing Assets}
\begin{tabular}{|p{4cm}|p{2cm}|p{2cm}|p{2cm}|}
\hline
\textbf{Asset Name} & \textbf{Status} & \textbf{Used This Session} & \textbf{Notes} \\
\hline
& & & \\
& & & \\
& & & \\
\hline
\end{tabular}

\subsection{Follower Updates}\index{Appendices!Session Log!Followers}

\subsubsection{Active Followers}\index{Appendices!Session Log!Active Followers}
\begin{tabular}{|p{3cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{2cm}|}
\hline
\textbf{Name} & \textbf{Role} & \textbf{Harm} & \textbf{Exposure} & \textbf{Status} \\
\hline
& & & & \\
& & & & \\
& & & & \\
\hline
\end{tabular}

\subsubsection{New Followers Recruited}\index{Appendices!Session Log!New Followers}
\begin{itemize}
    \item \textbf{Follower Name}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
    \item \textbf{Cost}: \_\_\_\_\_ XP
    \item \textbf{Downtime}: \_\_\_\_\_ days
    \item \textbf{Capabilities}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\end{itemize}

\subsection{Bond Development}\index{Appendices!Session Log!Bonds}

\subsubsection{Bonds Strengthened}\index{Appendices!Session Log!Strengthened Bonds}
\begin{itemize}
    \item \textbf{Character/NPC}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
    \item \textbf{Development}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\end{itemize}

\subsubsection{New Bonds Formed}\index{Appendices!Session Log!New Bonds}
\begin{itemize}
    \item \textbf{Character/NPC}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
    \item \textbf{Nature of Bond}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\end{itemize}

\subsection{Complication Resolution}\index{Appendices!Session Log!Complications}

\subsubsection{Complications Cleared}\index{Appendices!Session Log!Cleared Complications}
\begin{tabular}{|p{5cm}|p{5cm}|}
\hline
\textbf{Complication} & \textbf{Resolution Method} \\
\hline
& \\
& \\
& \\
\hline
\end{tabular}

\subsubsection{Ongoing Complications}\index{Appendices!Session Log!Ongoing Complications}
\begin{itemize}
    \item \textbf{Active Complications}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
    \item \textbf{CP Debt This Session}: +\_\_\_\_\_ CP per scene
\end{itemize}

\subsection{Boon Economy}\index{Appendices!Session Log!Boons}

\begin{itemize}
    \item \textbf{Boons Held}: \_\_\_\_\_ / 5 maximum
    \item \textbf{Boons Converted to XP}: \_\_\_\_\_ (2 Boons = 1 XP, max 2 XP/session)
    \item \textbf{Boons Spent on Assets}: \_\_\_\_\_
    \item \textbf{Boons from Bond Actions}: \_\_\_\_\_
\end{itemize}

\subsection{Session Rewards Summary}\index{Appendices!Session Log!Rewards}

\begin{tabular}{|p{5cm}|c|}
\hline
\textbf{Reward Type} & \textbf{XP Gained} \\
\hline
Table Attendance & +2 XP \\
Major Objective Reached & +\_\_\_\_\_ XP \\
Discovery or Lore Unlocked & +\_\_\_\_\_ XP \\
Hard Choice Embraced & +\_\_\_\_\_ XP \\
Complication Spotlight & +\_\_\_\_\_ XP \\
Bond/Flag Driven Play & +\_\_\_\_\_ XP \\
GM Curveball Award & +\_\_\_\_\_ XP \\
Complication Dividend (Cards) & +\_\_\_\_\_ XP \\
\textbf{Total Session XP} & \textbf{+\_\_\_\_\_ XP} \\
\hline
\end{tabular}

\subsection{Downtime Activities}\index{Appendices!Session Log!Downtime}

\begin{itemize}
    \item \textbf{Total Downtime Days}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
    \item \textbf{Activities Undertaken}:
    \begin{itemize}
        \item \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
        \item \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
        \item \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
    \end{itemize}
    \item \textbf{Haste Clock Status (if applicable)}: \_\_\_\_\_/\_\_\_\_\_ segments
\end{itemize}

\subsection{Campaign Clock Updates}\index{Appendices!Session Log!Campaign Clocks}

\begin{itemize}
    \item \textbf{Mandate}: \_\_\_\_\_ / 6
    \item \textbf{Crisis}: \_\_\_\_\_ / 6
    \item \textbf{Notable Changes}: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\end{itemize}

\subsection{Notes and Reflections}\index{Appendices!Session Log!Notes}

\vspace{3cm}

\end{section} 

\chapter*{Voices}\index{Voices}
\addcontentsline{toc}{chapter}{Voices}

\section*{Acasia}
\textbf{Elite.}
\begin{quote}
"I kept the ledgers when the last governors fled. Law is a map; ours burned. In Acasia the only borders left are curses you learn too late. The ink runs thin where the old roads used to be, and the names of places shift like smoke in the wind. I've seen three different cartographers try to chart our borders—each map disagreed with the last, and all of them were wrong. The curse doesn't care about parchment—it cares about blood, about the weight of oaths broken and promises forgotten. Every milepost whispers a different name now, and if you listen too long, you start to forget your own."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"You don't take shelter—you bargain with whatever lives in the dark. Pay in trinkets, or it takes memories. I'm light on both. Last week I traded my grandmother's wedding ring for a night's peace in an abandoned barn. The thing that took it looked almost human, but its eyes were wrong—like staring into deep water. Now I can't remember her face, only that she had kind hands. The ring's gone, the memory's gone, and I'm still hungry. There's a merchant two valleys over who'll give you a vial of forget-me-not tea for a copper piece—brew it right and you won't remember trading away your nightmares. Problem is, you also won't remember your children's names."
\end{quote}

\section*{Aelaerem}
\textbf{Elite.}
\begin{quote}
"Our hedges are not for show; they are promises plaited green. We tithe milk and song so the old neighbors stay on their side of dusk. Every spring we walk the boundaries, weaving new growth with old wood, singing the songs that hold the wild things at bay. The hedge-witch taught me that a properly tended hedge remembers every word spoken near it—every blessing, every curse, every secret whispered in the dark. We plant according to the moon's phase: rowan for protection, hawthorn for luck, and always, always, a sprig of rosemary to remember the dead. The old neighbors are particular about their due—too little song and they scratch at the roots; too much milk and they send the crows to peck your crops. Balance is everything."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"The flock startles before the wind turns. When they refuse the east meadow, I refuse it too. Hoofprints don't lie—even the small, barefoot ones. Last harvest moon, all the sheep faced inward, huddled together like they were afraid of something coming from the edges of the field. I found prints in the mud the next morning—too small for any animal I'd ever seen, too deep for the weight they should have been. The old shepherd said once that sheep see what we can't, that their fear is cleaner than ours. They know when the hedge-witch has been walking in her other shape, when the mummers have taken off their masks and forgotten to put them back on. I follow their lead because they've never led me wrong yet."
\end{quote}

\section*{Aeler}
\textbf{Elite.}
\begin{quote}
"Profit is a kind of ballast; it keeps a city steady when the deep knocks. We sell guarantees—tunnel, treaty, inheritance. Every vault has its price, every passage its fee, every word carved in stone its weight in silver. The Guild Princes don't rule from above—we rule from the ledgers, from the careful balance of what goes down and what comes up. When the mountain groans in the night, we don't pray—we calculate. How much stone will shift? What tunnels will need shoring? Which families will lose their birthright to the deep? The answers are always in the numbers, always in the careful accounting of risk versus reward. A guarantee properly worded is worth more than a crown, and lasts longer than most marriages."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"You learn to hear the stone thinking. If it dreams hungry, you wake the alarm and hope the lifts aren't already coming up full. The deep has moods like a sleeping giant—sometimes it snores, sometimes it turns over, sometimes it wakes up hungry. I've been down there for twelve years, and I can tell by the way the walls breathe whether we're safe or not. Last winter, the stone started singing in the third descent—a low hum that made the lamp-wicks flutter. The engineer said it was just gas pockets, but I knew better. Something was listening down there, something that remembered when the vaults were new and the guarantees meant something. We sealed that tunnel for a season, but the singing followed us up."
\end{quote}

\section*{Aelinnel}
\textbf{Elite.}
\begin{quote}
"Equations are cages for impossible beasts. Mind your bars; the numbers will bite if the proof has gaps. In the Mathlands, we don't just calculate—we domesticate. Every formula is a leash, every theorem a collar. I've seen what happens when a proof breaks—when the bars of an equation snap and the beast inside gets loose. Last solstice, the Geometer-General tried to prove the nature of time itself. The equation held for three days, then... something slipped through. Time doubled back on itself for a mile around the academy. Students aged fifty years in an hour, then un-aged just as quickly. The Geometer-General is still trying to solve for what got out—he won't eat, won't sleep, just stares at his chalkboard and mutters about prime numbers and the weight of infinity."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Master says the kettle is safe because it only screams on prime days. I marked my calendar with bandages. The first month I was apprenticed, I thought he was joking. Then the kettle started screaming on the 2nd, the 3rd, the 5th... always prime numbers. Master's got twenty-seven kettles in his workshop, and they all sing different songs. The copper one hums lullabies to keep the equations docile. The iron one growls warnings when the proofs start to strain. The silver kettle—that one's the worst. It only screams on Mersenne primes, and when it does, the numbers start moving on their own. I've seen equations crawl across the floor like spilled ink, looking for something to calculate. Master says that's what happens when you try to cage infinity—it fights back."
\end{quote}

\section*{Black Banner}
\textbf{Elite.}
\begin{quote}
"A contract is a country small enough to carry. We march inside its borders until someone pays to redraw them. Every clause is a territory, every signature a citizen. I've seen generals surrender entire provinces for the right wording in a rider, and kings give up daughters for a properly witnessed amendment. The beauty of a contract is that it's always negotiable—until the blood starts flowing. Then it becomes something else entirely. My company's banner has changed hands seventeen times in the last decade. Each new captain brings their own interpretation of the original charter, their own understanding of what we're contracted to protect. The men don't care who signs the papers—they care who pays their wages and who leads them into battle."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Uniform by day, coin by night. If the banners dance with Ykrul or Vilikari, I keep step and count my fingers after. Last campaign, we switched sides three times in as many weeks. First we were protecting a merchant convoy from river pirates, then we were the pirates attacking a noble's estate, then we were the noble's guards putting down a peasant uprising. The captain said it was all according to contract—that the wording was flexible enough to cover any interpretation. By the end, I couldn't remember whose coin I was spending or whose blood was on my sword. The only constant was the paymaster's seal, and even that looked different every payday."
\end{quote}

\section*{Ecktoria}
\textbf{Elite.}
\begin{quote}
"Empire is a habit; we still set tables as if guests will arrive crowned. What we lack in legions we spend in letters. The marble forums echo with the ghosts of senators who never existed, and the basilicas still ring with the voices of laws that were never passed. I've spent three years drafting a treaty with a kingdom that fell to plague twenty years ago—their ambassador arrives every spring, all ceremony and empty titles, and we discuss trade routes that lead to abandoned cities. The coin-house still mints coins bearing the face of an emperor who died before his coronation, because to stop would be to admit that the empire is smaller than we pretend. We are a nation of forms, of the careful ritual of governance performed for an audience of shadows."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"I stamp more seals than I eat meals. Paper walks farther than soldiers do now. Every document that crosses my desk needs three different stamps, two signatures, and a witness who swears they saw it written in triplicate. I've seen trade agreements that were longer than epic poems, marriage contracts that required their own libraries, and death certificates that needed to be notarized by the dead man's living relatives. The clerks have developed a kind of poetry to the process—we can tell just by looking at a document how important it is, how many departments it will need to visit before it's properly dead. The most tragic are the love letters that come through the system—by the time they're stamped and sealed, the lovers have usually forgotten why they wrote them."
\end{quote}

\section*{Kahfagia}
\textbf{Elite.}
\begin{quote}
"Currents are taxes paid in advance. We collect where winds consent and governors remember who taught them the tide. The sea is the greatest empire builder—every wave carries tribute, every storm redistributes wealth. I've seen merchant fleets founder on reefs that weren't there last season, and fishing villages prosper from currents that bring silver fish instead of the usual bait. The Admiralty Prefect doesn't rule from a throne—he rules from a chart room, watching the dance of wind and water that determines who prospers and who drowns. The Oshiiran taught us that the sea remembers everything—every dropped anchor, every spilled cargo, every prayer whispered to placate the deep. We collect not just the obvious tariffs, but the subtle ones: the price of safe passage, the cost of favorable winds, the fee for not mentioning the wrecks we know about."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Salt is honest work—your palms tell the truth by dusk. The empire is ropes and knots; I know which ones hold. Every morning I check the rigging on three different ships, and every evening I can tell by the calluses on my hands which lines were properly secured and which were just for show. The salt gets into everything—your clothes, your food, your dreams. I've been at sea for fifteen years, and I still wake up thinking I hear the waves. The old salts say that's how you know you're truly part of the fleet—when the sea starts talking to you in your sleep. The Vilikari officers think they understand navigation, but they're still learning to read the wind. They measure distance in leagues; we measure it in the way the spray tastes and the angle of the sun on the water."
\end{quote}

\section*{Linns}
\textbf{Elite.}
\begin{quote}
"Oars decide what laws can reach. My ships are courts that sit wherever the tide allows. The law is a fluid thing among the skerries—it changes with the depth of water, with the strength of the wind, with the phase of the moon. I've held court on ice floes, in fog banks, and once on a whale's back when the sea was calm enough to make it work. The River Jarls understand that justice isn't about permanence—it's about adaptation. A ruling that works in deep water might be useless in the shallows, and a law that serves the mainland is often meaningless among the islands. My authority flows like the tide—strong in some channels, weak in others, always moving, always changing. The only constant is the need for a strong arm and a steady hand on the tiller."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"We read the Yloka by taste—mud for shallows, iron for towns. When the rowlocks sing, a door is about to open. The river has its own language, and if you listen with more than your ears, it tells you everything you need to know. Sweet water means we're approaching a village that's been good to travelers. Bitter water warns of bandits or bad blood ahead. When the current changes direction suddenly, it's because someone upstream has made a decision that's going to affect us all. The old oarsmen say the river remembers every drop of blood spilled in its waters, and when the time is right, it brings the reckoning to the surface. I've seen fish rise with arrowheads in their bellies, and reeds grow around coins that were lost generations ago."
\end{quote}

\section*{Mistlands}
\textbf{Elite.}
\begin{quote}
"We count winters, not kings. The mountain signs the treaties; we only initial them. In the Mistlands, power doesn't come from crowns or charters—it comes from endurance. I've seen kingdoms rise and fall while our flocks grazed the same hills, watched empires crumble while our bells tolled the same hours. The mountain speaks in geological time, in the slow settling of stone and the patient accumulation of snow. Kings come and go like seasons, but the mountain remembers every treaty, every boundary stone, every drop of blood spilled on its slopes. We are the mountain's memory keepers, the ones who ensure that the old agreements hold even when the signatories have forgotten what they promised. The Dwarves understand this—they know that some things are meant to last longer than human ambition."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Mist keeps secrets. I keep sheep. Between us, we've hidden whole wars from the world. The mist comes down thick some mornings, so thick you can't see your own hands in front of your face. That's when the real work gets done—when the sheep move through the fog like ghosts, carrying messages in their wool, hiding refugees in their folds. I've guided armies through the mist without the enemy knowing they were there, smuggled nobles across borders that should have been impassable. The sheep don't care about politics—they just want good grass and clean water. But they're better at keeping secrets than most people, and they never betray a trust. The mist respects that kind of loyalty."
\end{quote}

\section*{Silkstrand}
\textbf{Elite.}
\begin{quote}
"Water is the city's memory; every debt finds its level. Our bravos keep the surface calm while old things move below. The canals remember everything—the weight of every boat, the cargo of every barge, the blood that's been spilled in their depths. I've seen merchants who thought they could escape their obligations by taking ship, only to find that the water carried their debts faster than their sails. The old things beneath the city are patient—they wait for the right moment to surface, when the weight of accumulated sins finally breaks the surface tension. We maintain the delicate balance between order and chaos, between the prosperity that flows on the surface and the corruption that festers below. Every bridge is a negotiation, every canal a compromise with forces that predate the city itself."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Oar in one hand, knife in the other. If the fog tastes bitter, don't sing—listen. The gondoliers know the city's secrets because we navigate by more than sight. The taste of the water tells us what's happening in the districts we can't see, the sound of our oars against the current warns of trouble ahead. When the fog comes down thick over the canals, it's not just visibility that's lost—communication breaks down, alliances shift, and old enemies find themselves rowing side by side. I've carried messages that were worth more than gold, and passengers who were worth less than the air they breathed. The fog doesn't care about your status—it just wants to know what you're really made of."
\end{quote}

\section*{Theona}
\textbf{Elite.}
\begin{quote}
"Hospitality binds sharper than iron. We keep the old table set, and our guests behave like saints—most nights. The Three Greens understand that a properly offered meal is a kind of magic—it creates obligations that last longer than any contract. I've seen wars averted because of a shared supper, alliances forged over a pot of properly blessed stew. The old table in the High Hall has seen more treaties signed over its surface than any council chamber in the mainland kingdoms. The guests know the rules—they eat what's offered, drink what's poured, and leave a gift that's appropriate to their station. Break the hospitality, and you break something fundamental about the way the world works. The Green Neighbors are particular about their due."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Lights on the water mean one of two things: a catch or a bargain. Bring extra salt either way. The fishing boats work by different rules than the mainland vessels—they understand that the sea is not just a source of food, but a partner in commerce. Every haul from the nets might contain something that's not quite fish, something that's been sent as payment for debts incurred in previous generations. The salt we use isn't just for preserving the catch—it's for seasoning the deals we make with forces that don't always follow human logic. I've seen nets come up heavy with silver coins mixed in with the herring, and I've seen them come up empty when the price wasn't right. The sea keeps its own ledgers, and it's not always interested in human currency."
\end{quote}

\section*{Thepyrgos}
\textbf{Elite.}
\begin{quote}
"Dispute is our sport; we wrestle with premises and shake truth out by the ankle. In the City of a Thousand Stairs, every argument is a climb, every conclusion a summit worth reaching. I've seen philosophers spend years debating the nature of a single word, climbing higher and higher on the scaffolding of logic until they're dizzy with the view. The Synod Hall echoes with the voices of scholars who've dedicated their lives to proving that the opposite of everything is also true. We don't just debate ideas—we inhabit them, live inside their implications until we understand their weight and texture. Truth in Thepyrgos is not a destination but a journey, and the most interesting truths are the ones that lead us to unexpected vistas."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"When the scholars argue, we add steps. Every answer needs somewhere noble to stand. The stonecutters and masons of Thepyrgos understand that knowledge has weight—it needs to be properly supported, given a foundation that won't shift under the pressure of new ideas. I've built lecture halls for philosophers who needed to pace while they thought, constructed viewing platforms for astronomers who wanted to get closer to the sky, and carved steps into solid rock for debates that were expected to last for months. The scholars pay well for quality work, but they also expect their surroundings to reflect the importance of their thoughts. A poorly constructed staircase can undermine an entire philosophical system—form and function are inseparable in our city."
\end{quote}

\section*{Ubral}
\textbf{Elite.}
\begin{quote}
"We own what we can defend at dawn. Feuds are calendars the dead write on our doors. In the highlands, property is not measured in acres but in the strength of your arm and the sharpness of your blade. I've seen territories change hands three times in a single day—won at sunrise, lost at noon, reclaimed by sunset. The clan chiefs understand that ownership is temporary, but honor is eternal. Every feud adds another name to the list, another obligation that will be passed down through generations. The dead write their calendars in blood, and the living are expected to honor their appointments. We don't sign treaties—we make promises that are witnessed by the stones and remembered by the wind."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Cattle know the safe passes better than chiefs do. I follow their noses and keep my knife friendly. The herders understand that the mountains have their own wisdom—the cattle can smell danger on the wind, sense the shifting of stone before it happens, know which paths are watched by unfriendly eyes. I've learned to read their behavior like a map, to understand that when they bunch together and face a particular direction, there's something worth paying attention to over the next ridge. The chiefs make their plans in council halls, but the cattle make the real decisions about where we're safe and where we're not. My knife is for protection, but it's also a tool for making offerings to the spirits who watch over the high passes."
\end{quote}

\section*{Valewood}
\textbf{Elite.}
\begin{quote}
"What grows too tall forgets its roots. Stones that dream of permanence become flowers. We are the forest's gardeners, not its jailers. The old empire sought to write its name in stone, but stone is deaf to whispers. The wood listens. It heard their boots, their swords, their screams—and then it dreamed them into mulch. What is buried becomes soil. What is forgotten becomes rain. What is remembered becomes new growth. The forest does not hate the past; it digests it. You cannot impose will on a tree that has forgotten more than you will ever know."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Silence is the forest's first language. Words are its second. When you speak thanks, you name a debt. When you name a debt, you invite the forest to collect. I have walked paths that opened for me like flower petals, and paths that closed behind me like fists. The difference was never in what I said, but in what I did not say. The forest counts your footfalls, not your gratitude. It remembers the weight of your need, not the sound of your praise. Offer something that cannot be named—your breath, your shadow, the space between your heartbeats. The wood accepts only what you do not realize you have given."
\end{quote}

\section*{Vhasia}
\textbf{Elite.}
\begin{quote}
"We dine from borrowed plates and plot at borrowed maps. The crown fits best when worn in rumor. In the Fractured Sun, legitimacy is not inherited but acquired—gathered like loose coins, picked up from the tables of fallen houses. I've seen pretenders who commanded more respect than the so-called legitimate heirs, dukes who ruled from cellars while their castles were occupied by strangers. The old court is scattered to the winds, but its forms and ceremonies live on in the dreams of those who would restore it. We are a nation of shadows and echoes, of rulers who govern from hiding and nobles who sign their names in code. The crown is not a symbol of authority—it's a target, and everyone knows it."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Dukes change faster than bowstrings. I keep both dry and swear as required. The countryside is a patchwork of competing authorities—this village owes fealty to a merchant who bought the title, that town answers to a priest who claims divine right, and the next hamlet follows a bandit chief who's proven he can protect them better than anyone else. I've served under seven different lords in as many months, and I've learned that loyalty is a luxury I can't afford. My bowstring stays dry because I never know when I might need to use it, and my oaths are carefully worded to leave room for interpretation. The only constant is change, and the only certainty is that today's master will be tomorrow's memory."
\end{quote}

\section*{Vilikari}
\textbf{Elite.}
\begin{quote}
"We trade with yesterday's enemy and sow where tomorrow's treaty will land. Soil forgives what banners won't. The Marches understand that enmity is seasonal—what burns today might nourish tomorrow's harvest. I've signed peace treaties with men whose fathers I helped bury, and declared war on allies whose grandfathers were my drinking companions. The soil doesn't care about the color of your banner—it only wants to know if you're bringing nutrients or poison. The Foedus Stone has seen more treaties carved into its surface than any other monument in the known world, and still the work continues. We are farmers of diplomacy, understanding that the best agreements are the ones that grow naturally from the conditions of the land."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"I've buried seed and brothers in the same field. Both come up when it rains. The frontier farmers understand that death and life are partners in the same dance—the bones of the fallen enrich the soil that feeds the living. I've seen graves that produced the finest wheat in the district, and battlefields that grew flowers so beautiful they attracted pilgrims from across the sea. The soil remembers everything, and it's not particular about the source of its nutrients. My brothers died defending this ground, and now it feeds my children. The cycle continues, and the land is patient. The Ykrul understand this—they know that conquest without cultivation is just temporary occupation."
\end{quote}

\section*{Viterra}
\textbf{Elite.}
\begin{quote}
"We rule by bridges. Break a span and watch titles wash away. In the Last Kingdom, power flows like water—controlled by the careful management of crossings and the strategic placement of obstacles. I've seen earls who commanded vast territories on paper lose everything when their bridges were destroyed, and minor nobles who controlled a single ford become the most important people in the region. The rivers are the arteries of Viterra, and those who control the flow of traffic across them control the destiny of the realm. Every bridge is a statement of authority, every ford a potential chokepoint. The Queen understands that governance is not about commanding obedience but about making compliance the easiest path available."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Coins talk louder than heralds at the ford. I hear who's winning before the town does. The ferry operators and bridge toll collectors are the true intelligence network of Viterra—we know who's paying what to whom, and we can read the political winds by the weight of the purses that cross our waters. I've seen merchants who could afford to pay double the normal toll but chose to haggle over a few coppers, and I've seen nobles who tried to cross without payment and found themselves waiting for weeks. The coins tell the story of the realm more accurately than any official report. When the payments start changing hands, when new faces appear at the fords, that's when you know the balance of power is shifting."
\end{quote}

\section*{Ways Between}
\textbf{Elite.}
\begin{quote}
"A threshold is a promise with hinges. I oil them so the dead don't squeal when they pass. The Gate-Wardens understand that every crossing is a negotiation, every passage a compact between the traveler and the forces that govern the spaces between spaces. I've seen doors that led to yesterday, windows that opened onto tomorrow, and thresholds that connected to places that never were and never will be. The hinges are the key—they determine not just whether a door opens, but what kind of opening it creates. A well-maintained threshold can be the difference between salvation and damnation, between arrival and endless wandering. The dead are particular about their passage—they remember every sound, every hesitation, every moment of doubt."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"Walk until your name sounds strange in your own mouth; that's how you know the road is listening. The pilgrims and wayfarers of the Spiritways understand that identity is fluid in the spaces between worlds. I've walked paths that made me forget my own face, crossed bridges that changed my name with every step, and navigated mazes that rearranged my memories like cards in a dealer's hands. The road doesn't just test your endurance—it tests your willingness to become someone new. When your name starts to feel foreign, when your past seems like someone else's story, that's when the path reveals its true nature. Most people never reach that point—they turn back when the road starts asking questions they're not prepared to answer."
\end{quote}

\section*{Ykrul}
\textbf{Elite.}
\begin{quote}
"Grass is a sea that drowns empires. We ride the waves and choose who to pull ashore. The Violet Steppe understands that power is not about holding territory but about controlling movement—about being able to appear anywhere and disappear just as quickly. I've seen great armies bogged down in grass that seemed harmless, and small warbands that could move faster than the wind across terrain that should have been impassable. The grass remembers every hoofbeat, every campfire, every drop of blood spilled on its surface. We are not conquerors but navigators, reading the signs that others miss and following paths that exist only in the memory of the steppe. Empires build walls; we build speed."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"If the wind smells of iron, veer north. If it smells of bread, saddle light—we won't be welcome. The scouts and outriders understand that the steppe speaks in scents and sounds that civilized people never learn to interpret. Iron means battle—recent or impending—and the smart scout knows when to avoid the area. Bread means settlement, and settlements mean complications. The Ykrul prefer the open grass where the only laws are the ones written in hoofprints and the only currency is the ability to move faster than your enemies. I've smelled smoke from fires that were fifty miles away, heard the whisper of horse hooves across ground that should have muffled all sound, and felt the approach of storms that wouldn't show on any civilized weather vane."
\end{quote}

\section*{Zakov}
\textbf{Elite.}
\begin{quote}
"Customs is a conversation, not a wall. We speak every dialect of gray. The Salt and Serpent understand that the most profitable trade happens in the spaces between legality and illegality, in the gray areas where definitions become flexible and boundaries shift according to the needs of the moment. I've negotiated deals that were technically legal but morally bankrupt, and arranged transactions that were illegal but socially necessary. The key is understanding that every law has its price, every regulation its exception. The Syndicate Princes don't rule through force—they rule through the careful manipulation of definitions, the strategic application of technicalities, and the patient cultivation of relationships with those who enforce the rules."
\end{quote}
\textbf{Commoner.}
\begin{quote}
"If a crate walks by itself, don't see it. If a coin falls alone, catch it for the man behind you. The dock rats and runners understand that Zakov operates on levels that the surface world never sees. I've watched cargo move itself from ship to warehouse without human hands, seen coins change ownership in ways that defied explanation, and learned that some of the most valuable goods in the city are the ones that don't appear on any manifest. The city has its own intelligence network—not the official kind with guards and couriers, but the kind that operates through whispers in dark alleys and signals that only the initiated understand. The smart runner knows when to look away and when to pay attention, when to speak and when to remain silent."
\end{quote}

\end{chapter}
