\documentclass[11pt]{article}

% --- Preamble (readable, lightweight) ---
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{hidelinks}

% --- Badge styles (WCAG-friendly foreground/background) ---
\definecolor{firmbg}{HTML}{E6F4EA}     % light green
\definecolor{firmfg}{HTML}{0B7A29}
\definecolor{probbg}{HTML}{E7F1FD}     % light blue
\definecolor{probfg}{HTML}{0B57D0}
\definecolor{appxbg}{HTML}{FFF4E5}     % light amber
\definecolor{appxfg}{HTML}{8C4A00}
\definecolor{pastbg}{HTML}{F1F3F4}     % light gray
\definecolor{pastfg}{HTML}{3C4043}

\newcommand{\badge}[3]{%
  {\setlength{\fboxsep}{2pt}\colorbox{#1}{\textcolor{#2}{\sffamily\footnotesize\textbf{#3}}}}%
}
\newcommand{\firm}{\badge{firmbg}{firmfg}{FIRM}}
\newcommand{\prob}{\badge{probbg}{probfg}{PROB}}
\newcommand{\appx}{\badge{appxbg}{appxfg}{APPROX}}
\newcommand{\past}{\badge{pastbg}{pastfg}{SOMETIME IN THE PAST}}

% --- List spacing ---
\setlist[itemize]{leftmargin=*, itemsep=4pt, topsep=4pt}
\setlist[description]{leftmargin=2.5cm, style=nextline, itemsep=6pt, topsep=6pt}

\title{In-World Timeline of \emph{Kon'reh} Lore}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Dating Schema}
\begin{description}
  \item[\firm] \textbf{Exact year attested} in-text or by cross-checked colophon.
  \item[\prob] \textbf{Probable range} inferred tightly from adjacent dated events.
  \item[\appx] \textbf{Approximate (circa)} decade/century estimate.
  \item[\past] \textbf{Intentionally undated;} known only to predate a later copy or mention.
\end{description}

\section*{Known \& Measured History}
\begin{itemize}
  \item \firm\ \textbf{223 AR} --- Testimony of \textbf{Legate Varro Pellanus}. Earliest narrative of a high-stakes, cross-cultural match under Ykrul duress.
  \item \firm\ \textbf{538 AR} --- \textbf{Aqyl of Thepyrgos} compiles the \emph{Corpus Canré}.
  \item \firm\ \textbf{553 AR} --- \textbf{Edict of Colonia Victrix} bans the \emph{Corpus}; formal study fractures.
  \item \firm\ \textbf{804 AR} --- \textbf{Safiya al--Fhara}, \emph{Diwān of Roads and Boards} surveys living traditions post-ban.
  \item \firm\ \textbf{856 AR} --- \textbf{Braedon Fenwood}, \emph{Concordance} compiled (editorial drive toward a ``cleaner book'').
  \item \firm\ \textbf{862 AR} --- \textbf{Kon'reh: Rules \& Core Lore Compendium} published; ``the book is done.''
  \item \firm\ \textbf{876 AR} --- \textbf{Markus Fenwood} enrolls at the University of Thepyrgos; studies under \textbf{Aqyl, son of Aqyl}.
  \item \firm\ \textbf{889 AR} --- \textbf{Letter from Alayse Fenwood} (explicitly dated ``Alayse, 889 AR'').
\end{itemize}

\section*{Ambiguous \& Purposeful Mystery}
\begin{itemize}
  \item \past\ \textbf{The Burning at Ash--Fenn}. Eyewitness account later copied (648 AR) describes the Witness bringing a trembling leaf to ``Aqyl, much aged,'' with ``his son---also Aqyl'' present. The event \emph{predates} the 648 copy; the year is deliberately withheld.
  \item \prob\ \textbf{The Aqyl Lineage}. ``Aqyl, son of Aqyl'' may be a direct descendant of the 538 compiler, a transmitted title within a learned line, or a literary choice emphasizing continuity; the texts preserve the ambiguity on purpose.
  \item \prob\ \textbf{The Witness \& The Ninth}. Practices (e.g., \emph{Salt Stitch}, \emph{Veil of Names}) are recorded with consequences (the Ash--Fenn dossier) while ontology remains unresolved; puzzles alter what readers notice, not tournament legality.
\end{itemize}

\section*{Style Notes (for future entries)}
\begin{itemize}
  \item \textbf{Order:} Year first; bold the key noun; then a one-line gloss.
  \item \textbf{Provenance clause:} After an em-dash, state where the date comes from (e.g., ``dated leaf,'' ``editorial colophon,'' ``copied in 648 AR'').
  \item \textbf{Ranges:} Use ``\textbf{c.\,630--650 AR}'' for \appx; use ``\textbf{by 648 AR}'' when anchored only by a later copy (\past).
  \item \textbf{Ambiguity verbs:} Prefer \emph{appears, is attested, is copied, is attributed} over definitive \emph{is/was} when evidence is indirect.
  \item \textbf{Do not upgrade uncertainty:} If a page does not date it, resist back-solving unless two independent anchors agree.
\end{itemize}

\end{document}
