%=====================================================================
%  Fate's Edge Character Sheet - Cantor
%=====================================================================
\documentclass[10pt]{article}
\usepackage[margin=1.5cm,landscape]{geometry}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage[most]{tcolorbox}

\definecolor{edgered}{HTML}{8C1E1E}
\definecolor{edgepaper}{HTML}{FAF5EB}
\definecolor{edgedark}{HTML}{1E1E28}

\pagecolor{edgepaper}
\color{edgedark}

\newtcolorbox{charbox}[2][]{%
    colback=edgepaper,
    colframe=edgered,
    fonttitle=\bfseries\sffamily,
    title=#2,
    sharp corners,
    boxrule=0.5pt,
    left=2pt,
    right=2pt,
    top=2pt,
    bottom=2pt,
    #1
}

\pagestyle{empty}

\begin{document}

\begin{center}
    {\Huge\bfseries\sffamily\color{edgered} Fate's Edge}\\[0.2em]
    {\large\sffamily Cantor — The Touched Voice}\\[0.5em]
    \rule{0.5\textwidth}{0.5pt}\\[1em]
\end{center}

\begin{minipage}[t]{0.48\linewidth}
    \begin{charbox}{Identity}
        \begin{tabular}{p{2.5cm} p{5cm}}
            \textbf{Name:} & \underline{\hspace{4cm}} \\
            \textbf{Heritage:} & \underline{\hspace{4cm}} \\
            \textbf{Background:} & \underline{\hspace{4cm}} \\
            \textbf{Patron:} & \underline{\hspace{4cm}} \\
            \textbf{Tier / XP:} & \underline{\hspace{2cm}} / \underline{\hspace{2cm}} \\
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Attributes \& Resources}
        \begin{tabular}{>{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm} >{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm}}
            \toprule
            \textbf{Attribute} & \textbf{Rating} & \textbf{Resource} & \textbf{Max/Track} \\
            \midrule
            Body & \underline{\hspace{1.5cm}} & Fatigue & \underline{\hspace{1.5cm}} / \underline{\hspace{1.5cm}} \\
            Wits & \underline{\hspace{1.5cm}} & Harm & \underline{\hspace{1.5cm}} / 3 \\
            Spirit & \underline{\hspace{1.5cm}} & Boons & \underline{\hspace{1.5cm}} / 5 \\
            Presence & \underline{\hspace{1.5cm}} & XP & \underline{\hspace{1.5cm}} \\
            \bottomrule
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Corruption Clock}
        \textbf{Clock Size = Spirit} \underline{\hspace{1.5cm}}\\
        \begin{tabular}{|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
            \hline
            1 & 2 & 3 & 4 & 5 & 6 \\
            \hline
        \end{tabular}\\
        \textbf{Mark Corruption when:} Pushing a Song, performing a Resonant Rite, or GM spends SB on occult activity.\\
        \textbf{When Clock fills:} Gain boon + burden from last Patron. Reset to Tier (min 1).\\
        \textbf{Current Corruption:} \underline{\hspace{1cm}} / \underline{\hspace{1.5cm}}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Repertoire}
        \textbf{Repertoire Clock [6]}\\
        \begin{tabular}{|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
            \hline
            1 & 2 & 3 & 4 & 5 & 6 \\
            \hline
        \end{tabular}\\
        At 2: -1 DV to Songs (min 2). At 4: +1 die Performance. At 6: Learn one Standard Rite as Song.
    \end{charbox}
\end{minipage}
\hfill
\begin{minipage}[t]{0.48\linewidth}
    \begin{charbox}{Songs \& Performance}
        \textbf{Sing a Song:} 1 action, resolves next turn. Roll Lore + Performance vs DV.\\
        \textbf{Push It:} Resolve immediately. Mark 1 Fatigue and mark Corruption. GM gains 1 SB.\\
        \textbf{High Cantor (18 XP):} Sing Standard Rites as Songs; resolve instantly.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Cantor Talents}
        \textbf{Resonant Performance (3 XP):} +1 die when performing for 5+ audience.\\
        \textbf{Song Weaver (4 XP):} Combine two compatible Songs for +1 Effect.\\
        \textbf{Steady Voice (2 XP):} When Pushing, may mark +1 Fatigue instead of Corruption (once/scene).\\
        \textbf{Patron’s Dedication (4 XP):} Bind to single Patron. Singing another’s Song worsens Position.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Patrons of the Voice}
        \textbf{Palinode (Queen of Encores):} Hymn Against Dread, The Single True Note, Crownpiece.\\
        \textbf{Thrysos (King of Revels):} The Overflowing Cup, The Horned Dance, Endless Feast.\\
        \textbf{Livaea (Crimson Courtier):} Velvet Whisper, Shared Cup, Binding Vow (social focus).\\
        \textbf{Malachai (Cruel Messenger):} Honeyed Curse, False Dawn, The Messenger’s Burden.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Roleplay Hooks}
        \underline{\hspace{7cm}}\\
        \underline{\hspace{7cm}}
    \end{charbox}
\end{minipage}

\vspace{1em}
\begin{center}
    {\small\sffamily\itshape “You think you need a lute? My larynx is older than any tree. Hum, and the world will listen.” — The Grey Wanderer}
\end{center}

\end{document}
