%=====================================================================
%  Fate's Edge Character Sheet − Invoker
%=====================================================================
\documentclass[10pt]{article}
\usepackage[margin=1.5cm,landscape]{geometry}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage[most]{tcolorbox}

\definecolor{edgered}{HTML}{8C1E1E}
\definecolor{edgepaper}{HTML}{FAF5EB}
\definecolor{edgedark}{HTML}{1E1E28}

\pagecolor{edgepaper}
\color{edgedark}

≠wtcolorbox{charbox}[2][]{%
    colback=edgepaper,
    colframe=edgered,
    fonttitle=\bfseries\sffamily,
    title=#2,
    sharp corners,
    boxrule=0.5pt,
    left=2pt,
    right=2pt,
    top=2pt,
    bottom=2pt,
    #1
}

\pagestyle{empty}

\begin{document}

\begin{center}
    {\Huge\bfseries\sffamily\color{edgered} Fate's Edge}\\[0.2em]
    {łarge\sffamily Invoker — The Gambler at the Crossroads}\\[0.5em]
    \rule{0.5\textwidth}{0.5pt}\\[1em]
\end{center}

\begin{minipage}[t]{0.48łinewidth}
    \begin{charbox}{Identity}
        \begin{tabular}{p{2.5cm} p{5cm}}
            \textbf{Name:} & ℎ̲pace{4cm}} \\
            \textbf{Heritage:} & ℎ̲pace{4cm}} \\
            \textbf{Background:} & ℎ̲pace{4cm}} \\
            \textbf{Tier / XP:} & ℎ̲pace{2cm}} / ℎ̲pace{2cm}} \\
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Attributes & Resources}
        \begin{tabular}{>{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm} >{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm}}
            ⊤rule
            \textbf{Attribute} & \textbf{Rating} & \textbf{Resource} & \textbf{Max/Track} \\
            ∣rule
            Body & ℎ̲pace{1.5cm}} & Fatigue & ℎ̲pace{1.5cm}} / ℎ̲pace{1.5cm}} \\
            Wits & ℎ̲pace{1.5cm}} & Harm & ℎ̲pace{1.5cm}} / 3 \\
            Spirit & ℎ̲pace{1.5cm}} & Boons & ℎ̲pace{1.5cm}} / 5 \\
            Presence & ℎ̲pace{1.5cm}} & XP & ℎ̲pace{1.5cm}} \\
            ⊥tomrule
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Symbols & Obligation}
        \begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
            ℎline
            \textbf{Patron} & \textbf{Symbol} & \textbf{Obligation} & \textbf{Status} \\
            ℎline
            1: & ℎ̲pace{1.5cm}} & ℎ̲pace{1cm}} & ℎ̲pace{1.5cm}} \\
            ℎline
            2: & ℎ̲pace{1.5cm}} & ℎ̲pace{1cm}} & ℎ̲pace{1.5cm}} \\
            ℎline
            3: & ℎ̲pace{1.5cm}} & ℎ̲pace{1cm}} & ℎ̲pace{1.5cm}} \\
            ℎline
            4: & ℎ̲pace{1.5cm}} & ℎ̲pace{1cm}} & ℎ̲pace{1.5cm}} \\
            ℎline
        \end{tabular}\\
        \textbf{Total Obligation:} ℎ̲pace{1cm}} / ℎ̲pace{1.5cm}} (Spirit + Presence)\\
        \textbf{Overflow:} Each segment over Capacity = 1 Fatigue.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Ritual Invocation}
        \textbf{Time:} Significant Time (DV in rounds). +1 round per DV.\\
        \textbf{Cost:} Mark the Rite’s Obligation.\\
        \textbf{Symbol Benefit:} If Symbol openly displayed, reduce DV by 1 and Obligation by 1 (min 0).\\
        \textbf{Crack the Seal:} Resolve instantly. Set Symbol to Compromised, mark double Obligation (min +2).
    \end{charbox}
\end{minipage}
ℎfill
\begin{minipage}[t]{0.48łinewidth}
    \begin{charbox}{Invoker Talents}
        \textbf{Borrowed Grace (Free):} Spend 1 Boon or 1 Fatigue to gain +1 die to a single action. Mark +1 Obligation.\\
        \textbf{Crack Specialist (6 XP):} Reduce Obligation cost of Crack the Seal by 1.\\
        \textbf{Symbol Resonance (4 XP):} +1 die to ritual roll when using a Symbol.\\
        \textbf{Ritual Mastery (4 XP):} Reduce backlash SB by 1; first SB each scene becomes your Boon instead.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Sample Invoker Patrons}
        \textbf{Morag the Hag (Hidden Costs):} The Speaking Knot, Threefold Exchange, Crossroads Court.\\
        \textbf{The Confessor (Burden & Exchange):} Echo of Sin, Burden Transference, The Cauldron’s Secret.\\
        \textbf{The Ninth (Hyperknowledge):} Ninth Figure, Hypercognitive Lattice, The Ninth Revelation.\\
        \textbf{Ikasha (Shadow):} Touch the Umbral Veil, Umbral Reservoir, Become the Shadow.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Rites Prepared}
        ℎ̲pace{7cm}}\\
        ℎ̲pace{7cm}}\\
        ℎ̲pace{7cm}}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Roleplay Hooks}
        ℎ̲pace{7cm}}\\
        ℎ̲pace{7cm}}
    \end{charbox}
\end{minipage}

\vspace{1em}
\begin{center}
    {\small\sffamilyℹtshape “I do not serve. I rent. And when the landlord comes to collect, I had best have the coin ready.” — The Grey Wanderer}
\end{center}

\end{document}
