%=====================================================================
%  Fate's Edge Character Sheet − Runekeeper
%=====================================================================
\documentclass[10pt]{article}
\usepackage[margin=1.5cm,landscape]{geometry}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage[most]{tcolorbox}

\definecolor{edgered}{HTML}{8C1E1E}
\definecolor{edgepaper}{HTML}{FAF5EB}
\definecolor{edgedark}{HTML}{1E1E28}

\pagecolor{edgepaper}
\color{edgedark}

≠wtcolorbox{charbox}[2][]{%
    colback=edgepaper,
    colframe=edgered,
    fonttitle=\bfseries\sffamily,
    title=#2,
    sharp corners,
    boxrule=0.5pt,
    left=2pt,
    right=2pt,
    top=2pt,
    bottom=2pt,
    #1
}

\pagestyle{empty}

\begin{document}

\begin{center}
    {\Huge\bfseries\sffamily\color{edgered} Fate's Edge}\\[0.2em]
    {łarge\sffamily Runekeeper — Agent of the Inked Covenant}\\[0.5em]
    \rule{0.5\textwidth}{0.5pt}\\[1em]
\end{center}

\begin{minipage}[t]{0.48łinewidth}
    \begin{charbox}{Identity}
        \begin{tabular}{p{2.5cm} p{5cm}}
            \textbf{Name:} & ℎ̲pace{4cm}} \\
            \textbf{Heritage:} & ℎ̲pace{4cm}} \\
            \textbf{Background:} & ℎ̲pace{4cm}} \\
            \textbf{Patron:} & ℎ̲pace{4cm}} \\
            \textbf{Tier / XP:} & ℎ̲pace{2cm}} / ℎ̲pace{2cm}} \\
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Attributes & Resources}
        \begin{tabular}{>{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm} >{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm}}
            ⊤rule
            \textbf{Attribute} & \textbf{Rating} & \textbf{Resource} & \textbf{Max/Track} \\
            ∣rule
            Body & ℎ̲pace{1.5cm}} & Fatigue & ℎ̲pace{1.5cm}} / ℎ̲pace{1.5cm}} \\
            Wits & ℎ̲pace{1.5cm}} & Harm & ℎ̲pace{1.5cm}} / 3 \\
            Spirit & ℎ̲pace{1.5cm}} & Boons & ℎ̲pace{1.5cm}} / 5 \\
            Presence & ℎ̲pace{1.5cm}} & XP & ℎ̲pace{1.5cm}} \\
            ⊥tomrule
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Obligation Tracker}
        \textbf{Capacity:} Spirit ℎ̲pace{1cm}} + Presence ℎ̲pace{1cm}} = ℎ̲pace{1.5cm}}\\
        \textbf{Current Obligation:} ℎ̲pace{1cm}} / ℎ̲pace{1.5cm}}\\
        \begin{tabular}{|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
            ℎline
            1 & 2 & 3 & 4 & 5 & 6 \\
            ℎline
            μlticolumn{6}{c}{ℎ̲pace{6cm}}} \\
        \end{tabular}\\
        \textbf{Patron’s Name:} ℎ̲pace{4cm}}\\
        \textbf{Obligation Overflow:} Each segment over Capacity = 1 Fatigue.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Codex & Thiasos}
        \textbf{Codex:} ℎ̲pace{6cm}}\\
        (Your book of Rites. If lost or Compromised, cannot invoke Rites.)\\
        \textbf{Thiasos (Familiar):} ℎ̲pace{6cm}}\\
        (Your Patron’s fragment. If Compromised, −1 die to all Rites.)
    \end{charbox}
\end{minipage}
ℎfill
\begin{minipage}[t]{0.48łinewidth}
    \begin{charbox}{Rites Known}
        \textbf{Low Rites (Tier 1, Obligation 4–5):}\\
        ℎ̲pace{7cm}}\\
        ℎ̲pace{7cm}}\\
        \textbf{Standard Rites (Tier 2, Obligation 6–8):}\\
        ℎ̲pace{7cm}}\\
        ℎ̲pace{7cm}}\\
        \textbf{High Rites (Tier 3, Obligation 7–8):}\\
        ℎ̲pace{7cm}}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Patron’s Gift & Push It}
        \textbf{Patron’s Gift (Imbuement):} Once/scene as action, touch an item. For the scene, +1 die to thematic skill, and first use ignores 1 Story Beat. Mark +1 Obligation.\\
        \textbf{Push It:} Once/scene, mark +1 Obligation to amplify a Rite (+1 die or +1 Effect).
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Signature Talent}
        \textbf{Patron’s Dedication (4 XP):} Bind your voice to a single Patron. Your Corruption timer is always keyed to that Patron. Singing another Patron’s Song worsens Position by one step.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Runekeeper Talents}
        \textbf{Efficient Invocation (4 XP):} Reduce Obligation cost of one specific Low Rite by 1 (min 0).\\
        \textbf{Patron’s Favor (8 XP):} Once/session, ignore Obligation cost for a Rite.\\
        \textbf{Runebound Resilience (18 XP):} Convert up to 2 Harm into Obligation instead of damage.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Roleplay Hooks}
        ℎ̲pace{7cm}}\\
        ℎ̲pace{7cm}}
    \end{charbox}
\end{minipage}

\vspace{1em}
\begin{center}
    {\small\sffamilyℹtshape “I do not beg. I record. And in the recording, I become the hand that writes the law.” — The Grey Wanderer}
\end{center}

\end{document}
