%=====================================================================
%  Fate's Edge Character Sheet − Runic Warrior (Melee Runekeeper)
%=====================================================================
\documentclass[10pt]{article}
\usepackage[margin=1.5cm,landscape]{geometry}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage[most]{tcolorbox}

\definecolor{edgered}{HTML}{8C1E1E}
\definecolor{edgepaper}{HTML}{FAF5EB}
\definecolor{edgedark}{HTML}{1E1E28}

\pagecolor{edgepaper}
\color{edgedark}

≠wtcolorbox{charbox}[2][]{%
    colback=edgepaper,
    colframe=edgered,
    fonttitle=\bfseries\sffamily,
    title=#2,
    sharp corners,
    boxrule=0.5pt,
    left=2pt,
    right=2pt,
    top=2pt,
    bottom=2pt,
    #1
}

\pagestyle{empty}

\begin{document}

\begin{center}
    {\Huge\bfseries\sffamily\color{edgered} Fate's Edge}\\[0.2em]
    {łarge\sffamily Runic Warrior — Melee Runekeeper}\\[0.5em]
    \rule{0.5\textwidth}{0.5pt}\\[1em]
\end{center}

\begin{minipage}[t]{0.48łinewidth}
    \begin{charbox}{Identity}
        \begin{tabular}{p{2.5cm} p{5cm}}
            \textbf{Name:} & ℎ̲pace{4cm}} \\
            \textbf{Heritage:} & ℎ̲pace{4cm}} \\
            \textbf{Background:} & ℎ̲pace{4cm}} \\
            \textbf{Patron:} & ℎ̲pace{4cm}} \\
            \textbf{Tier / XP:} & ℎ̲pace{2cm}} / ℎ̲pace{2cm}} \\
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Attributes & Combat Resources}
        \begin{tabular}{>{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm} >{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm}}
            ⊤rule
            \textbf{Attribute} & \textbf{Rating} & \textbf{Resource} & \textbf{Max/Track} \\
            ∣rule
            Body & ℎ̲pace{1.5cm}} & Fatigue & ℎ̲pace{1.5cm}} / ℎ̲pace{1.5cm}} \\
            Wits & ℎ̲pace{1.5cm}} & Harm & ℎ̲pace{1.5cm}} / 3 \\
            Spirit & ℎ̲pace{1.5cm}} & Boons & ℎ̲pace{1.5cm}} / 5 \\
            Presence & ℎ̲pace{1.5cm}} & XP & ℎ̲pace{1.5cm}} \\
            ⊥tomrule
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Obligation Tracker}
        \textbf{Capacity:} Spirit ℎ̲pace{1cm}} + Presence ℎ̲pace{1cm}} = ℎ̲pace{1.5cm}}\\
        \textbf{Current Obligation:} ℎ̲pace{1cm}} / ℎ̲pace{1.5cm}}\\
        \begin{tabular}{|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
            ℎline
            1 & 2 & 3 & 4 & 5 & 6 \\
            ℎline
        \end{tabular}\\
        \textbf{Patron:} ℎ̲pace{4cm}}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Combat Rites & Imbuements}
        \textbf{Patron’s Gift:} Once/scene, imbue weapon. For the scene, +1 die to melee and +1 thematic. Mark +1 Obligation.\\
        \textbf{Radiant Smite (Standard, 9 XP):} Next melee strike flares with dawnfire. Upgrade Effect, +1 Harm, or force 1 SB.\\
        \textbf{Thornveil (Standard, 7 XP):} Raise a living barrier. Enemies suffer −1 die to approach.\\
        \textbf{Covenant Blaze (High, 13 XP):} All sworn within are haloed; aggressors suffer −1 die.
    \end{charbox}
\end{minipage}
ℎfill
\begin{minipage}[t]{0.48łinewidth}
    \begin{charbox}{Runic Warrior Talents}
        \textbf{Spirit Shield (6 XP):} When you Defend, convert first Harm 1 to Fatigue.\\
        \textbf{Battlefield Mastery (6 XP):} When outnumbered, +1 Effect and enemies −1 die for 3 exchanges.\\
        \textbf{Mystic Bulwark (8 XP):} When you Defend, shield an adjacent ally. Spend 2 Fatigue to extend to all in Near.\\
        \textbf{Effortless Charge (4 XP):} Once/scene, ignore Fatigue cost of Charge action.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Weapon & Armor}
        \begin{tabular}{>{\bfseries}p{3cm} p{4.5cm}}
            ⊤rule
            \textbf{Equipment} & \textbf{Stats} \\
            ∣rule
            Primary Weapon: & ℎ̲pace{4cm}} \\
            Armor: & ℎ̲pace{4cm}} \\
            Shield: & ℎ̲pace{4cm}} \\
            ⊥tomrule
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Patron Archetypes}
        \textbf{Mykkiel (Paladin):} Final Admonition, smite foes, consecrate ground.\\
        \textbf{Grimmir (Druid):} Thornveil, speak with beasts, heal blight.\\
        \textbf{Oath of Flame (Crusader):} Radiant Smite, Lay on Hands, vow−based defense.\\
        \textbf{Khemesh (Warlock):} Maw Opens, crushing pressure, sanity damage.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Roleplay Hooks}
        ℎ̲pace{7cm}}\\
        ℎ̲pace{7cm}}
    \end{charbox}
\end{minipage}

\vspace{1em}
\begin{center}
    {\small\sffamilyℹtshape “The Codex is not merely a tool but a contract made visible, a ledger of debts paid and owed.” — The Grey Wanderer}
\end{center}

\end{document}
