%=====================================================================
%  Fate's Edge Character Sheet - Witch
%=====================================================================
\documentclass[10pt]{article}
\usepackage[margin=1.5cm,landscape]{geometry}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage[most]{tcolorbox}

\definecolor{edgered}{HTML}{8C1E1E}
\definecolor{edgepaper}{HTML}{FAF5EB}
\definecolor{edgedark}{HTML}{1E1E28}

\pagecolor{edgepaper}
\color{edgedark}

\newtcolorbox{charbox}[2][]{%
    colback=edgepaper,
    colframe=edgered,
    fonttitle=\bfseries\sffamily,
    title=#2,
    sharp corners,
    boxrule=0.5pt,
    left=2pt,
    right=2pt,
    top=2pt,
    bottom=2pt,
    #1
}

\pagestyle{empty}

\begin{document}

\begin{center}
    {\Huge\bfseries\sffamily\color{edgered} Fate's Edge}\\[0.2em]
    {\large\sffamily Witch — The Quiet Architect}\\[0.5em]
    \rule{0.5\textwidth}{0.5pt}\\[1em]
\end{center}

\begin{minipage}[t]{0.48\linewidth}
    \begin{charbox}{Identity}
        \begin{tabular}{p{2.5cm} p{5cm}}
            \textbf{Name:} & \underline{\hspace{4cm}} \\
            \textbf{Heritage:} & \underline{\hspace{4cm}} \\
            \textbf{Background:} & \underline{\hspace{4cm}} \\
            \textbf{Patron:} & \underline{\hspace{4cm}} \\
            \textbf{Tier / XP:} & \underline{\hspace{2cm}} / \underline{\hspace{2cm}} \\
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Attributes \& Resources}
        \begin{tabular}{>{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm} >{\bfseries}p{2.5cm} >{\bfseries}p{1.5cm}}
            \toprule
            \textbf{Attribute} & \textbf{Rating} & \textbf{Resource} & \textbf{Max/Track} \\
            \midrule
            Body & \underline{\hspace{1.5cm}} & Fatigue & \underline{\hspace{1.5cm}} / \underline{\hspace{1.5cm}} \\
            Wits & \underline{\hspace{1.5cm}} & Harm & \underline{\hspace{1.5cm}} / 3 \\
            Spirit & \underline{\hspace{1.5cm}} & Boons & \underline{\hspace{1.5cm}} / 5 \\
            Presence & \underline{\hspace{1.5cm}} & XP & \underline{\hspace{1.5cm}} \\
            \bottomrule
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Promise Timers}
        \textbf{Promise Timer [4] — Tracks a debt or obligation:}\\
        \begin{tabular}{|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
            \hline
            1 & 2 & 3 & 4 \\
            \hline
        \end{tabular}\\
        \textbf{Promise Timer [6]:}\\
        \begin{tabular}{|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
            \hline
            1 & 2 & 3 & 4 & 5 & 6 \\
            \hline
        \end{tabular}
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{The Three Layers}
        \textbf{Echo:} Past memory. Work with knots and cords.\\
        \textbf{Veil:} Present boundary. Work with salt and chalk.\\
        \textbf{Flow:} Direction of change. Work with omission and the Ninth.
    \end{charbox}
\end{minipage}
\hfill
\begin{minipage}[t]{0.48\linewidth}
    \begin{charbox}{Hedge Gifts (No Roll)}
        \textbf{Steady Hand:} Once/scene, touch wounded → remove 1 Fatigue.\\
        \textbf{Warm Hand:} Once/scene, touch → next physical action +1 die (comfort) or -1 die (warning).\\
        \textbf{Hearth-Sense:} Once/scene, ask GM yes/no about a threshold or debt.\\
        \textbf{Salt Line:} Once/scene, pour salt line → spirits must test Spirit+Resolve to cross.\\
        \textbf{The Unlit Candle:} Once/session, extinguish light source → dim/dark zone.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Quick Workings (Roll)}
        Name a threshold, choose a layer, roll Wits+Lore or Spirit+Insight. Effect: door sticks, candle gutters, pursuer slips.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Witch Patrons}
        \textbf{Morag the Hag:} Unspoken costs, faerie bargains. The Speaking Knot, Threefold Exchange.\\
        \textbf{Inaea (Angel of the Spider):} Webs of connection, mercy as control. Hearth-Thread Knot, Strand of Inevitability.\\
        \textbf{Ikasha (She Who Sleeps):} Thresholds, shadows, forgotten things. Touch the Umbral Veil.\\
        \textbf{Isoka (Serpent’s Shedding):} Transformation, venom. Venomous Benediction, The Shedding.
    \end{charbox}
    
    \vspace{0.5em}
    
    \begin{charbox}{Roleplay Hooks}
        \underline{\hspace{7cm}}\\
        \underline{\hspace{7cm}}
    \end{charbox}
\end{minipage}

\vspace{1em}
\begin{center}
    {\small\sffamily\itshape “The hedge is what keeps the wolves from the flock. I am the one who tends the hedge.” — The Grey Wanderer}
\end{center}

\end{document}
