\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{parskip}
\setlist{leftmargin=*,itemsep=0.25em}
\titleformat{\section}{\large\bfseries}{}{0em}{}[\vspace{-0.25em}\hrule\vspace{0.4em}]
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

\begin{document}
\begin{center}
{\LARGE Kon'reh — Quick Start}\\
\vspace{0.25em}
{\small Game of Apex, Sanctum, and Reforge}
\end{center}

\section*{Goal}
Capture your opponent’s \textbf{Blue} (Arbiter), then \textbf{prevent its return}. After Blue is captured, its owner has \textbf{five of their own turns} to \textbf{plant a banner} (end a move) on the enemy \textbf{Home Apex}. If they succeed, their Blue returns (see \emph{Reforge}). If they fail, \textbf{you win}. 

\section*{Turn Order}
Players alternate one move per turn. \textbf{Second player gets an opening double-move} (two different pieces) on their first turn.

\section*{Board, Pieces, Movement}
\begin{itemize}
  \item Board is an \textbf{8$\times$8 diamond}. Four corner \emph{Apexes}: Home, Opposing, two \emph{Sanctums}.
  \item Your army has four classes:
    \begin{itemize}
      \item \textbf{Blue (Arbiter)} — your only piece with specials (see below).
      \item \textbf{Orange} — long lane control.
      \item \textbf{Red} — short lane posts/screens.
      \item \textbf{Green} — \emph{spawned} runners via \emph{Seed}; not present at setup.
    \end{itemize}
  \item \textbf{Slides are lane-straight.} A move traverses \emph{one} lane (no turns in a single move).
  \item \textbf{Zone of Control (ZoC):} each piece projects ZoC to its four edge-adjacent squares. You may \emph{enter} enemy ZoC, but \textbf{entering ends the move}. You may not pass \emph{through} enemy ZoC.
  \item \textbf{Captures (non-Blue):} Red/Orange/Green capture by \textbf{displacement} (step onto enemy; remove it).
\end{itemize}

\section*{Blue (Arbiter) Specials}
On a Blue turn you may \textbf{slide, then do \emph{one} special}. Never special\,$\to$\,slide.
\begin{itemize}
  \item \textbf{Displacement (S:D):} step 1 along the lane onto an adjacent enemy (remove it).
  \item \textbf{Hop-capture (S:H):} jump over one adjacent enemy along the lane to the empty square beyond (remove the jumped piece; landing must be empty).
  \item \textbf{ZoC gate:} If your Blue’s slide \emph{entered} enemy ZoC this turn, you \emph{cannot} special afterward.
  \item \textbf{Crown Stagger:} if you use your \emph{second distinct} special in the same Blue life (e.g., H \emph{and} D), your Blue becomes \textbf{Rooted} until your next turn.
\end{itemize}

\section*{Twin Apex Seed (spawning Greens)}
If your Blue \textbf{ends a move on a Sanctum}, and the opposite Sanctum is empty, you may \textbf{Seed}: place a \textbf{Green} on the \emph{opposite} Sanctum, respecting a \textbf{global cap of 6 Greens} on the board.
\begin{itemize}
  \item \textbf{Rooted on Seed:} the Blue that Seeded is \textbf{Rooted} until your next turn.
  \item \textbf{Mobilization Delay:} a Blue cannot Seed on its \textbf{first departure from Home} in a given life.
  \item \textbf{Same-Sanctum ban (per life):} if a Blue \emph{returned by Reforge} was placed on a Sanctum, that Blue may \emph{not} Seed from that \emph{same} Sanctum for the rest of its \emph{current life}.
\end{itemize}

\section*{Central Four (the Cross)}
The 2$\times$2 center diamond.
\begin{itemize}
  \item \textbf{Stay cap:} a Blue may end at most \textbf{3 of its own turns} in the Cross per life.
  \item \textbf{Exclusion:} after a Blue leaves the Cross, it may not re-enter for the next \textbf{2 of its own turns}.
  \item \textbf{Practice rule:} only enter with at least one \emph{certified exit} next turn (XS\,$\ge$\,1).
\end{itemize}

\section*{Reforge (after Blue is captured)}
When your Blue is captured, you have \textbf{five of your own turns} to plant a banner on the enemy Home Apex.
\begin{itemize}
  \item \textbf{Plant succeeds:} remove the planting piece, then \textbf{return your Blue} by one of:
    \begin{itemize}
      \item \textbf{Opposing Apex (their Home):} pay \textbf{1 Green} (sacrifice).
      \item \textbf{Either Sanctum:} free, but \textbf{this Blue may not Seed from that same Sanctum} for the rest of this life.
      \item \textbf{Your Home Apex:} free.
    \end{itemize}
  \item \textbf{Refresh:} a reforged Blue returns with \textbf{both specials refreshed}.
  \item \textbf{Failure:} if you do not plant within five of your turns, you \textbf{lose immediately}.
\end{itemize}

\section*{State Tags (track at the table)}
Use small markers/dice; call them aloud.
\begin{itemize}
  \item \texttt{[CF: in $x$/3]} — Cross stay count for Blue’s current life.
  \item \texttt{[Excl: $y$]} — Cross re-entry exclusion turns remaining.
  \item \texttt{[S:H]} / \texttt{[S:D]} — Blue specials spent this life.
  \item \texttt{[Rooted]} — Blue cannot move until your next turn.
  \item \texttt{[RC $n$/5]} — Reforge countdown for the side whose Blue was captured.
  \item \texttt{[CapC A--B]} — captures tally (optional).
  \item \texttt{G6 Dial} — global Green count (0–6).
\end{itemize}

\section*{Recommended Learning Path}
\begin{enumerate}[label=\arabic*)]
  \item \textbf{Tutor games (no Seed).} Learn slides, ZoC stops, basic captures.
  \item \textbf{Add Blue specials.} Demonstrate H and D; warn about \emph{Stagger}.
  \item \textbf{Add Cross discipline.} Enter only with XS\,$\ge$\,1; respect 3-stay/2-exclusion.
  \item \textbf{Enable Seed.} Teach Mobilization Delay; introduce an \emph{SSI} idea: Seed only if punish is 2+ plies away or screened.
  \item \textbf{Reforge drill.} Play a capture and run the five-turn banner race once per session.
  \item \textbf{Full games with tags.} Use the state markers above until it’s second nature.
\end{enumerate}

\section*{Table Etiquette (fast heuristics)}
\begin{itemize}
  \item \textbf{Count exits before center.} One certified exit is playable; two is safe.
  \item \textbf{Don’t spend both cuts lightly.} If you must use H and D in one life, plan to Stagger when it’s safe.
  \item \textbf{Seed for tempo, not vanity.} If your Blue will be Rooted into a punish, wait.
  \item \textbf{Banner math beats material.} If you can force a failed Reforge, trade pieces freely.
\end{itemize}

\end{document}
