\section{Ikasha — She Who Sleeps}

\subsection{Domain and Epithets}

Ikasha is the hush between footfalls, the patience of dark water, the black-feathered watcher at every threshold. In stillness she gathers what might be, at crossroads she whispers of what may yet come. Ravens circle her, bearing secrets between worlds. Her followers learn to move unseen and speak unremembered, becoming shadows that slip between what is and what could be.

Ikasha does not act; she waits. She does not command; she suggests. Her power is not in force but in possibility—the hidden door, the forgotten name, the secret that, if spoken, would change everything. Those who serve her become keepers of thresholds, listeners at the edge of sleep, and witnesses to the shape of things not yet born.

\subsection{Cultural Origins}

Ikasha is honored wherever shadows pool and secrets gather:

\begin{itemize}
\item In the \textbf{Mistlands}, she is the Bell-Shrouded. Bell-wardens leave a single rope unknotted for her, a silent invitation.
\item In \textbf{Zakov}, she is the Tide Watcher, a smuggler's patron whose shrines are hidden in sea caves at low tide.
\item In \textbf{Aeler}, she is the Deep Listener. Miners whisper to her before descending into new shafts. The stone answers—or does not.
\item In \textbf{Ashaan}, she is the Unbound Chain, the secret patron of the Breath-less, invoked in root cellars and sealed rooms.
\end{itemize}

\subsection{Warrior Archetype: The Shadow Warrior}

A Runic Warrior of Ikasha is not a front-line combatant. They are a \textit{Shadow Warrior}—a ninja, a spy, an unseen blade who strikes once and vanishes. Their Codex is a shadow-cloak or a black feather. Their Thiasos is a raven-spirit that watches from rooftops, a shadow-cat that moves without sound, or a sentient blade that drinks light.

\subsection{Codex Form}

The Shadow Warrior's Codex is born of darkness and secrecy:

\begin{itemize}
\item \textbf{Shadow-Cloak:} A cloak of black wool or silk that seems to drink light. When still, the wearer is difficult to notice. When moving, the cloak trails like smoke.
\item \textbf{Raven's Quill:} A black feather that writes in invisible ink. The ink appears only when held over a flame—or when a secret is spoken nearby.
\item \textbf{Silence Bell:} A small bronze bell with no clapper. It makes no sound—but when it would have rung, the Shadow Warrior knows they have been noticed.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Raven-Spirit:} A black bird that remembers faces. It can identify any person it has seen before, and it knows when someone is lying. It never speaks above a whisper.
\item \textbf{Shadow-Cat:} A great cat made of darkness and stillness. It moves without sound. It can pass through a locked door if there is a gap—any gap—beneath it.
\item \textbf{The Raven's Quill (Sentient):} A feather that speaks in rustles. It can erase a single written word from a document once per session. The word is gone as if it had never been written.
\end{itemize}

\subsection{Core Mechanic: The Unseen}

When the Shadow Warrior is not observed (no witnesses within line of sight), they gain +1 die to all physical actions. The shadow is their ally.

When the Shadow Warrior attacks from a hidden position, they gain +1 Effect. The first strike is the only strike.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Unlit Candle (4 XP)}

\textbf{Tags:} [STEALTH], [DECEPTION], [SHADOW]

\textbf{Threshold:} A candle snuffed with a whisper.

You step into the space between notices. Gain +1 die to a single Stealth or Subterfuge roll, and if you succeed, you may also declare one minor detail "forgotten" by any observer (e.g., "they don't remember my face," "they left the key in the lock").

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Crossroads Raven (5 XP)}

\textbf{Tags:} [OMEN], [NAVIGATION], [SECRET]

\textbf{Threshold:} Three black feathers scattered at a crossing.

A raven appears, visible only to you and your allies. It offers an omen: you may reroll one failed navigation, pursuit, or escape roll this scene, or force one enemy to hesitate (lose their next minor action).

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Umbral Reservoir (8 XP)}

\textbf{Tags:} [SHADOW], [RESTORATION], [BOOST]

\textbf{Threshold:} Water gathered on a moonless night, held in a black vessel.

Draw on deep reserves of patience and shadow. Choose one: gain +2 dice on your next Stealth, Deception, or Resolve roll; clear 1 Fatigue from yourself or a touched ally; or gain one free success on an attempt to escape restraint or pursuit (no roll needed).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Unlatched Secret (9 XP)}

\textbf{Tags:} [SECRET], [TRUTH], [UNLOCK], [MIND]

\textbf{Threshold:} A key that has never been used, bent in a circle.

Choose one: persuade a mundane threshold to open (one lock, seal, or barrier opens as if never secured) or compel a truthful answer to one direct question about a secret the target knows (they resist with Resolve DV 4).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Umbral Dream (14 XP)}

\textbf{Tags:} [SHADOW], [DREAM], [FATE], [TELEPORT], [INVISIBLE]

\textbf{Threshold:} Stand in absolute darkness or at a deserted crossroads at midnight.

Become a living shadow or reach into latent potential. Choose one: shadow form (intangible to mundane harm, can pass through small gaps, +2 dice to Stealth) or whisper of possibility (reroll any single roll after seeing the result, before outcome is declared) or hidden path (bypass one major obstacle as if a secret way had always been there).

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Shadow Meld (6 XP)}

\textbf{Requirement:} Ikasha

Once per session as an action, step into any shadow larger than yourself and emerge from another shadow within Near range. You cannot be targeted while moving between shadows. Mark 1 Fatigue after using this talent.

\subsubsection{Unseen Strike (4 XP)}

\textbf{Requirement:} Ikasha

When you attack from a hidden position, your target suffers -1 die to defend. They do not see the threat until it is too late.

\subsubsection{Raven's Witness (8 XP)}

\textbf{Requirement:} Ikasha, Spirit 3+

A raven appears when you invoke this talent. It will observe a single target for one hour, then return to you. You may ask the GM one question about what the raven witnessed (e.g., "Did they meet anyone?" "Where did they hide the key?"). The raven cannot report speech—only actions and locations.

\subsection{Roleplaying a Shadow Warrior}

\begin{itemize}
\item You do not fight fair. You fight unseen, unheard, unforgettable only in the damage you leave behind.
\item You may be called to infiltrate, to spy, to deliver messages that must not be intercepted. Your shadow is your shield.
\item You may struggle with the weight of secrets. You know things that could ruin people. You keep them. The silence is heavy.
\item You may be asked to trust—to show your face, to speak your name, to step into the light. The shadow does not trust easily.
\end{itemize}

\subsection{Sample NPC: The Raven's Keeper of the Breath-less Cell}

\textbf{Archetype:} Shadow Warrior

\textbf{Patron:} Ikasha

\textbf{Codex:} A raven's quill that writes in invisible ink

\textbf{Thiasos:} A raven-spirit that watches from rooftops

\textbf{Personality:} Silent, watchful, patient. She has not spoken her true name in twenty years. She is not sure she remembers it.

\textbf{Conflict:} A secret she has kept for a decade is about to surface. The person it would ruin is innocent—of this crime, at least. The shadow says: hide it. The heart says: let it go. The raven watches.

\subsection{GM Guidance: Intrusions for Ikasha}

\begin{itemize}
\item A raven lands on the Shadow Warrior's shoulder and whispers a secret they never wanted to know. The secret is true.
\item The Shadow Warrior's shadow stretches into the shape of a woman with a crown of black feathers. It moves when they are still.
\item A door the Shadow Warrior passed through earlier is now locked from the other side—and the key is in their pocket. They did not put it there.
\item The Shadow Warrior dreams of a crossroads they have never seen. On the road, someone is waiting. They do not know who. They are not sure they want to find out.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
