\section{The Silent Choir — Restraint and Protective Omission}

\subsection{Domain and Epithets}

The Silent Choir is the gathered voice of truths wisely left unsaid, confessions granted mercy, and secrets kept to spare a life from breaking. It is not absence but presence: the practiced wisdom that some knowledge wounds more than it heals. Its keepers are wardens of confidence and shelterers of the vulnerable. They learn to carry burdens alone, to let volatile truths sleep, and to shield others from fires that revelation would feed.

Unlike the Witness, who reveals truth at any cost, or Ikasha, who hides secrets for power, the Silent Choir teaches restraint as protection. A secret is not a weapon; it is a weight. To speak it may free the speaker but shatter the listener. The Choir's followers judge not what is true, but what is kind.

\subsection{Cultural Origins}

The Silent Choir is honored wherever discretion is valued over disclosure:

\begin{itemize}
\item In \textbf{Thepyrgos}, the Silent Choir is invoked by archivists who seal dangerous texts behind a second lock and whisper the combination to no one.
\item In \textbf{Aelaerem}, the Choir is the Unspoken Guest, a presence at deathbeds who listens but never speaks. Its symbol is a finger pressed to sealed lips.
\item In \textbf{Lethai} courts, the Choir is the Silence Between Clauses, the pause in a contract where something important is deliberately omitted.
\item In \textbf{Vhasia}, the Choir is the Bell-Shrouded, invoked when a confession must be witnessed but never repeated.
\end{itemize}

\subsection{Warrior Archetype: The Silent Blade}

A Runic Warrior of the Silent Choir is not an assassin or a spy. They are a \textit{Silent Blade}—a warrior who has sworn an oath of restraint, who fights only when necessary, who protects the vulnerable and keeps the secrets of the dead. Their Codex is a sealed lips tattoo or a grey cord. Their Thiasos is a memory-spirit that records confessions, a hollowed bird that cannot speak, or a sentient blade that hums only when a secret is near.

\subsection{Codex Form}

The Silent Blade's Codex is a testament to restraint:

\begin{itemize}
\item \textbf{Sealed Lips Tattoo:} A tattoo of a mouth sewn shut with silver thread. It is visible on the back of the hand or the throat. It is a vow made visible.
\item \textbf{Grey Cord:} A braided cord of undyed wool, tied around the wrist. Each knot is a secret kept. The cord tightens when the wearer is tempted to speak.
\item \textbf{Silence Bell:} A small bronze bell with no clapper. It cannot ring. It is a reminder that not every truth needs a voice.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Memory-Spirit:} A spirit that cannot speak, but can replay sounds it has witnessed—a confession, a scream, a name spoken in the dark. It does not judge. It only remembers.
\item \textbf{Hollowed Bird:} A small grey bird with milk-white eyes. It cannot speak. It witnesses. It never blinks.
\item \textbf{The Sealed Lips (Sentient):} A mask of clay or porcelain, its mouth sewn shut with silver thread. It does not speak—it listens. It can recall any conversation it witnessed with perfect clarity.
\end{itemize}

\subsection{Core Mechanic: The Kept Secret}

When the Silent Blade keeps a secret that would harm someone if revealed, they gain 1 Boon. This Boon can only be used to protect, to heal, or to shield another from harm.

When the Silent Blade is asked to reveal a secret they have sworn to keep, they may test Resolve (DV 4) to refuse without penalty. On a failure, they mark 1 Fatigue—the weight of the refusal.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Gentle Seal (5 XP)}

\textbf{Tags:} [SILENCE], [PROTECT], [BOOST], [DEBT]

\textbf{Threshold:} A white feather or cloth that has drunk whispered confidences.

Choose one: gentle silence (hush a space with respectful quiet; all present gain +1 die to actions of discretion, confidentiality, or de-escalation this scene) or shielded truth (ward one piece of information from discovery; gain +2 dice to resist extraction or revelation of the protected truth).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Quiet Ear (4 XP)}

\textbf{Tags:} [SILENCE], [KNOWLEDGE], [PROTECT]

\textbf{Threshold:} A shell held to the ear, or a door held ajar.

You listen without being heard. For one scene, you may hear any conversation within Near range, even through doors or thin walls, as long as you remain still and silent. You cannot be detected by mundane listening.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Burden Bearer (8 XP)}

\textbf{Tags:} [DEBT], [TRANSFER], [HEAL], [WITNESS]

\textbf{Threshold:} A grey cord tied around your wrist.

Name a person crushed by guilt, grief, or a secret they cannot bear. For one scene, they ignore all penalties from that burden. They are not free—they are carried. If that person then uses this reprieve to harm another, the burden returns doubled.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Unspoken Vow (7 XP)}

\textbf{Tags:} [VOW], [BOND], [SILENCE], [WITNESS]

\textbf{Threshold:} A sealed envelope or a locked box containing a written promise.

Swear an oath that will not be spoken aloud. The oath is witnessed by the Silent Choir itself. Breaking the oath causes the oath-breaker to suffer -1 die to all social rolls for one week—not from judgment, but from the weight of their own silence broken.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Unburdened Confession (13 XP)}

\textbf{Tags:} [DEBT], [HEAL], [TRUTH], [WITNESS]

\textbf{Threshold:} A grey cord, a candle, and a willing creature.

Absorb a single traumatic memory or secret from a willing creature, erasing it from their mind permanently. You gain the memory as your own (with all emotional weight), and the creature is freed from its effects. You gain 1 permanent Corruption and mark +2 Obligation. You cannot use this rite again until you have spent a downtime scene in meditation or confession.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Burden Steady (4 XP)}

\textbf{Requirement:} The Silent Choir

You may carry one additional psychological Condition without stacking penalties. The weight is not less—you have simply learned to bear it.

\subsubsection{Calm Presence (6 XP)}

\textbf{Requirement:} The Silent Choir, Spirit 3+

Once per scene, when you touch a creature suffering from Fear, Guilt, or Despair, you may remove that Condition. The creature clears 1 Fatigue if the Condition was particularly heavy.

\subsubsection{Final Vow (10 XP)}

\textbf{Requirement:} The Silent Choir, Tier III+

Once per campaign, you may take a dying person's last secret and carry it permanently. You gain +1 permanent Corruption, but you may never reveal that secret—not even under magical compulsion. The secret becomes part of you. The Silent Choir witnesses.

\subsection{Roleplaying a Silent Blade}

\begin{itemize}
\item You do not speak of what you know. You carry it. The weight is yours to bear.
\item You may be called to witness confessions, to protect the vulnerable from their own truths, or to silence those who would speak what should remain unsaid.
\item You may struggle with the burden. Every secret is a stone. The stones pile. The back bends. The Silent Choir does not offer relief—only purpose.
\item You may be asked to speak—to reveal a truth that would save a life, to break a vow for a greater good. The Choir does not forbid speech. It asks only that you be certain.
\end{itemize}

\subsection{Sample NPC: Sister Mira of the Sealed Lips}

\textbf{Archetype:} Silent Blade

\textbf{Patron:} The Silent Choir

\textbf{Codex:} A grey cord tied around her wrist, each knot a secret kept

\textbf{Thiasos:} A hollowed bird that witnesses but never speaks

\textbf{Personality:} Quiet, steady, inexhaustibly patient. She has not spoken a word in seven years. She does not need to. Her silence says everything.

\textbf{Conflict:} A secret she has carried for thirty years is killing her—slowly, quietly, one sleepless night at a time. The secret is her own. The Silent Choir does not forbid self-confession. But she has not yet found the courage. The bird watches. The cord tightens.

\subsection{GM Guidance: Intrusions for The Silent Choir}

\begin{itemize}
\item A secret the Silent Blade thought buried resurfaces. Someone knows. They are not threatening—yet.
\item The Silent Blade's grey cord tightens suddenly. They have been tempted to speak. The cord remembers.
\item A dying person asks the Silent Blade to witness a confession. The confession is terrible. The Silent Blade must carry it.
\item The Silent Blade dreams of a mouth sewn shut—their own. When they wake, they cannot speak for one hour. The Choir is testing them.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
