\section{Aveh — The Rider Behind the Storm}

\subsection{Domain and Epithets}

Aveh is the faceless rider at the horizon's edge, neither man nor woman, both and neither. The Ykrul call them Aveh, a liminal spirit beyond hearth and vow, a storm-shadow that offers liberty at the cost of belonging. To travel with Aveh is to taste unchained freedom—and to risk being erased by the storm that follows. They are the patron of the outcast, the non-conforming, the genderless, and the forgotten. Aveh asks no oath and demands no loyalty, but every gift leaves a scar of absence.

Aveh does not command. They offer. The road is open; the storm is patient. Those who ride with them learn that freedom is not a destination—it is a constant leaving.

\subsection{Cultural Origins}

Aveh appears wherever boundaries are crossed and identities shed:

\begin{itemize}
\item In \textbf{Ykrul}, Aveh is a sky-spirit that rides with winter gales; shamans invoke Aveh to hide a dying person's tracks from the dead.
\item In \textbf{Linn}, Aveh is a sea-fog rider; sailors whisper Aveh's name to slip through blockades, trading a memory for safe passage.
\item In \textbf{Vilikari} marches, Aveh is the Unbannered, invoked by those who have left their clan and seek a new name.
\item In \textbf{Aeler}, Aveh is the Deep Erasure, prayed to by those who must disappear into the mountain's depths and never return.
\end{itemize}

\subsection{Warrior Archetype: The Unbound Blade}

A Runic Warrior of Aveh is not a soldier or a sentinel. They are an \textit{Unbound Blade}—a ronin, a wanderer, a warrior who has cut all ties and answers to no master. Their Codex is a storm-banner or a broken sword. Their Thiasos is a wind-spirit that carries their scent away, a horse that leaves no tracks, or a sentient blade that remembers every owner it has outlived.

\subsection{Codex Form}

The Unbound Blade's Codex is a testament to freedom:

\begin{itemize}
\item \textbf{Storm-Banner:} A torn flag on a broken pole. It twists in wind that no one else can feel. Planted in the ground, it creates a zone where pursuers suffer -1 die to track.
\item \textbf{Broken Sword:} A blade snapped at the hilt, carried as a reminder. Its edge is still sharp. Its purpose is unfinished.
\item \textbf{Erased Map:} A parchment map with all names scratched out. The roads remain. The destinations are blank. The traveler chooses.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Wind-Spirit:} A translucent figure of moving air. It carries the Runekeeper's scent away, confuses tracking animals, and whispers warnings in a voice like dry grass.
\item \textbf{Trackless Horse:} A grey horse that leaves no hoofprints. It cannot be followed by mundane means. It knows every road—and every way off the road.
\item \textbf{The Storm-Banner (Sentient):} A torn flag that speaks in the howl of wind. Planted, it creates a zone of forgetting; those within lose track of time and direction.
\end{itemize}

\subsection{Core Mechanic: The Unremembered}

When the Unbound Blade successfully escapes pursuit or hides their tracks, they gain 1 Boon. This Boon can only be used on Stealth, Deception, or Escape rolls.

When the Unbound Blade is forgotten—by a witness, a ledger, a pursuer—they may mark 1 Fatigue to gain +1 die on their next action. The forgetting is not a gift; it is a tool.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Unlatched Step (4 XP)}

\textbf{Tags:} [ESCAPE], [STEALTH], [FREEDOM]

\textbf{Threshold:} A breath cast into the wind, or a line drawn in dust.

Choose one: storm-step (slip free of notice or restraint; gain +1 die on one escape or evasion roll this scene) or exile's banner (mark a target as "outside"; the target suffers -1 die on all social rolls for the scene).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Forgotten Road (6 XP)}

\textbf{Tags:} [ERASURE], [AREA], [FREEDOM]

\textbf{Threshold:} A threshold or roadway under open sky, or stormwater rubbed into skin.

Choose one: erase the road (all attempts to follow you this session suffer -2 dice; in exchange, erase part of your own path—lose a detail, a memory, a name chosen by the GM) or rider's mark (a companion gains +2 dice to resist capture or coercion, but others forget them more easily).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Storm's Refuge (12 XP)}

\textbf{Tags:} [INVISIBLE], [AREA], [ERASURE], [FREEDOM]

\textbf{Threshold:} A banner or cloak raised in storm.

Your group is unseen or forgotten by pursuers for the scene. At the end, all present must mark 1 Corruption or forget one bond (their choice).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Unremembered Name (13 XP)}

\textbf{Tags:} [ERASURE], [TRANSFORM], [FATE], [FREEDOM]

\textbf{Threshold:} A circle of ash or dust drawn at a boundary, and a personal token.

For one week, your name is erased from mundane records and casual memory. Those who knew you must succeed on a Resolve test (DV 4) to recall you. You gain +2 dice to all rolls to avoid detection, impersonate another, or pass unseen.

\textbf{Obligation:} 3

\subsubsection{High Rite: Rite of the Horizon's Devour (15 XP)}

\textbf{Tags:} [ERASURE], [AREA], [FATE], [DEATH]

\textbf{Threshold:} A circle of ash or dust drawn at a boundary, and a storm that answers.

All thresholds in a zone collapse. Doors lead nowhere; borders dissolve; memory of the place becomes vague. In exchange, something of you is erased—your reflection, your name, or your presence in a specific person's memory (GM's choice).

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Storm-Step (6 XP)}

\textbf{Requirement:} Aveh

Once per session as a move action, vanish in a swirl of wind and dust and reappear anywhere within Near range that you can see. This movement does not provoke opportunity attacks. Mark 1 Fatigue after using this talent.

\subsubsection{Forgotten Face (4 XP)}

\textbf{Requirement:} Aveh

Witnesses suffer -2 dice to recall your description. You are not invisible—merely unmemorable. The effect lasts until you do something remarkable.

\subsubsection{Escape Clause (8 XP)}

\textbf{Requirement:} Aveh, Tier II+

Once per session, when you are cornered or capture is inevitable, you may declare an escape route that was not there before. The GM must accept it, but may impose a cost—you left something behind, or someone else paid the price for your freedom.

\subsection{Roleplaying an Unbound Blade}

\begin{itemize}
\item You do not swear oaths. You make promises. You keep them—or you don't. The road does not judge.
\item You may be called to guide fugitives, to erase evidence, or to help someone disappear. You know the roads that do not appear on any map.
\item You may struggle with the cost of freedom. Every escape leaves something behind. Every erasure takes a memory. The storm does not ask what you are losing.
\item You may be asked to stay—to make a home, to keep a name, to be remembered. Aveh does not forbid roots. But the storm does not nourish them.
\end{itemize}

\subsection{Sample NPC: The Dust-Walker of the Unbannered}

\textbf{Archetype:} Unbound Blade

\textbf{Patron:} Aveh

\textbf{Codex:} A storm-banner that twists in wind no one else feels

\textbf{Thiasos:} A wind-spirit that carries his scent away

\textbf{Personality:} Quiet, restless, unnamed. He has not used his true name in ten years. He is not sure he remembers it. The wind does not ask.

\textbf{Conflict:} Someone is looking for him—not an enemy, not a creditor. A child who claims to be his daughter. He does not remember her. The storm would say: run. The heart says: wait. The wind is still.

\subsection{GM Guidance: Intrusions for Aveh}

\begin{itemize}
\item A storm erases the Unbound Blade's tracks—and also the memory of their face from the last town they visited.
\item The Unbound Blade's shadow detaches and wanders off for an hour. It returns with dirt on its feet. It will not say where it went.
\item A forgotten debt suddenly resurfaces. Someone the Unbound Blade wronged now knows exactly where they are.
\item The Unbound Blade dreams of a crossroads where all roads lead to the same place—a door they have never seen. On the other side, someone is waiting. The storm does not say who.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
