\section{Rites of the Shadow (Ikasha, The Silent Choir, Aveh)}

These Rites are available to Runic Warriors of the shadow patrons—those who move unseen, who keep secrets, who walk between notice and forgetfulness. Each Rite is a tool of stealth, silence, and erasure.

\subsection{Ikasha — She Who Sleeps}

Ikasha's Rites favor patience, concealment, and the power of the unnoticed. Her followers do not fight fair—they fight from shadow, from stillness, from the space between heartbeats.

\subsubsection{Low Rite: Rite of the Unlit Candle (4 XP)}

\textbf{Tags:} [STEALTH], [DECEPTION], [SHADOW]

\textbf{Threshold:} A candle snuffed with a whisper.

You step into the space between notices. Gain +1 die to a single Stealth or Subterfuge roll, and if you succeed, you may also declare one minor detail "forgotten" by any observer (e.g., "they don't remember my face," "they left the key in the lock").

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Crossroads Raven (5 XP)}

\textbf{Tags:} [OMEN], [NAVIGATION], [SECRET]

\textbf{Threshold:} Three black feathers scattered at a crossing.

A raven appears, visible only to you and your allies. It offers an omen: you may reroll one failed navigation, pursuit, or escape roll this scene, or force one enemy to hesitate (lose their next minor action).

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Umbral Reservoir (8 XP)}

\textbf{Tags:} [SHADOW], [RESTORATION], [BOOST]

\textbf{Threshold:} Water gathered on a moonless night, held in a black vessel.

Draw on deep reserves of patience and shadow. Choose one: gain +2 dice on your next Stealth, Deception, or Resolve roll; clear 1 Fatigue from yourself or a touched ally; or gain one free success on an attempt to escape restraint or pursuit (no roll needed).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Unlatched Secret (9 XP)}

\textbf{Tags:} [SECRET], [TRUTH], [UNLOCK], [MIND]

\textbf{Threshold:} A key that has never been used, bent in a circle.

Choose one: \textit{unlatched door} (persuade a mundane threshold to open—one lock, seal, or barrier opens as if never secured; does not work on magical wards) or \textit{secret keeper} (compel a truthful answer to one direct question about a secret the target knows; they resist with Resolve DV 4).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Umbral Dream (14 XP)}

\textbf{Tags:} [SHADOW], [DREAM], [FATE], [TELEPORT], [INVISIBLE]

\textbf{Threshold:} Stand in absolute darkness or at a deserted crossroads at midnight.

Become a living shadow or reach into latent potential. Choose one: \textit{shadow form} (intangible to mundane harm, can pass through small gaps, +2 dice to Stealth for the scene), \textit{whisper of possibility} (reroll any single roll after seeing the result, before outcome is declared), or \textit{hidden path} (bypass one major obstacle as if a secret way had always been there—GM gains 1 SB as the world adjusts).

\textbf{Obligation:} 3

\subsection{The Silent Choir — Restraint and Protective Omission}

The Silent Choir's Rites favor discretion, burden-bearing, and the power of the unspoken. Their followers carry secrets so others do not have to.

\subsubsection{Low Rite: Rite of the Gentle Seal (5 XP)}

\textbf{Tags:} [SILENCE], [PROTECT], [BOOST], [DEBT]

\textbf{Threshold:} A white feather or cloth that has drunk whispered confidences.

Choose one: \textit{gentle silence} (hush a space with respectful quiet; all present gain +1 die to actions of discretion, confidentiality, or de-escalation this scene) or \textit{shielded truth} (ward one piece of information from discovery; gain +2 dice to resist extraction or revelation of the protected truth).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Quiet Ear (4 XP)}

\textbf{Tags:} [SILENCE], [KNOWLEDGE], [PROTECT]

\textbf{Threshold:} A shell held to the ear, or a door held ajar.

You listen without being heard. For one scene, you may hear any conversation within Near range, even through doors or thin walls, as long as you remain still and silent. You cannot be detected by mundane listening.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Burden Bearer (8 XP)}

\textbf{Tags:} [DEBT], [TRANSFER], [HEAL], [WITNESS]

\textbf{Threshold:} A grey cord tied around your wrist.

Name a person crushed by guilt, grief, or a secret they cannot bear. For one scene, they ignore all penalties from that burden. They are not free—they are carried. If that person then uses this reprieve to harm another, the burden returns doubled.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Unspoken Vow (7 XP)}

\textbf{Tags:} [VOW], [BOND], [SILENCE], [WITNESS]

\textbf{Threshold:} A sealed envelope or a locked box containing a written promise.

Swear an oath that will not be spoken aloud. The oath is witnessed by the Silent Choir itself. Breaking the oath causes the oath-breaker to suffer -1 die to all social rolls for one week—not from judgment, but from the weight of their own silence broken.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Unburdened Confession (13 XP)}

\textbf{Tags:} [DEBT], [HEAL], [TRUTH], [WITNESS]

\textbf{Threshold:} A grey cord, a candle, and a willing creature.

Absorb a single traumatic memory or secret from a willing creature, erasing it from their mind permanently. You gain the memory as your own (with all emotional weight), and the creature is freed from its effects. You gain 1 permanent Corruption and mark +2 Obligation. You cannot use this rite again until you have spent a downtime scene in meditation or confession.

\textbf{Obligation:} 3

\subsection{Aveh — The Rider Behind the Storm}

Aveh's Rites favor freedom, erasure, and the power of the forgotten. Their followers are unbound—they leave no tracks, no memories, no debts.

\subsubsection{Low Rite: Rite of the Unlatched Step (4 XP)}

\textbf{Tags:} [ESCAPE], [STEALTH], [FREEDOM]

\textbf{Threshold:} A breath cast into the wind, or a line drawn in dust.

Choose one: \textit{storm-step} (slip free of notice or restraint; gain +1 die on one escape or evasion roll this scene) or \textit{exile's banner} (mark a target as "outside"; the target suffers -1 die on all social rolls for the scene).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Forgotten Road (6 XP)}

\textbf{Tags:} [ERASURE], [AREA], [FREEDOM]

\textbf{Threshold:} A threshold or roadway under open sky, or stormwater rubbed into skin.

Choose one: \textit{erase the road} (all attempts to follow you this session suffer -2 dice; in exchange, erase part of your own path—lose a detail, a memory, a name chosen by the GM) or \textit{rider's mark} (a companion gains +2 dice to resist capture or coercion, but others forget them more easily).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Storm's Refuge (12 XP)}

\textbf{Tags:} [INVISIBLE], [AREA], [ERASURE], [FREEDOM]

\textbf{Threshold:} A banner or cloak raised in storm.

Your group is unseen or forgotten by pursuers for the scene. At the end, all present must mark 1 Corruption or forget one bond (their choice). The storm does not erase without cost.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Unremembered Name (13 XP)}

\textbf{Tags:} [ERASURE], [TRANSFORM], [FATE], [FREEDOM]

\textbf{Threshold:} A circle of ash or dust drawn at a boundary, and a personal token.

For one week, your name is erased from mundane records and casual memory. Those who knew you must succeed on a Resolve test (DV 4) to recall you. You gain +2 dice to all rolls to avoid detection, impersonate another, or pass unseen.

\textbf{Obligation:} 3

\subsubsection{High Rite: Rite of the Horizon's Devour (15 XP)}

\textbf{Tags:} [ERASURE], [AREA], [FATE], [DEATH]

\textbf{Threshold:} A circle of ash or dust drawn at a boundary, and a storm that answers.

All thresholds in a zone collapse. Doors lead nowhere; borders dissolve; memory of the place becomes vague. In exchange, something of you is erased—your reflection, your name, or your presence in a specific person's memory (GM's choice).

\textbf{Obligation:} 3

\subsection{Rite Synergies}

Rites from different shadow patrons can combine in powerful ways:

\begin{itemize}
\item \textbf{Unlit Candle + Gentle Seal:} You are unseen and unheard. You move through guarded spaces as if you were never there.
\item \textbf{Umbral Dream + Storm's Refuge:} Your group becomes intangible and forgotten. No one saw you enter. No one will remember you left.
\item \textbf{Burden Bearer + Unburdened Confession:} You carry the weight of others' secrets—and when the weight is too great, you may release it into the Cord. The spiral continues.
\item \textbf{Unlatched Step + Forgotten Road:} You escape and leave no trace. Pursuers cannot follow—and soon cannot remember why they were pursuing.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
