\chapter{Common Failures and Their Lessons}

A ledger of mistakes I have made or witnessed. Each failure taught me something the Weave does not put in textbooks.

\section{The Candle That Would Not Die}

A novice attempted to extinguish a candle with [SILENCE]. He reasoned that sound and flame were connected. The Weave disagreed. The candle did not go out. It became \emph{perfectly silent}—no crackle, no hiss, no pop—while burning twice as bright. The student learned: tags affect exactly what they name. [SILENCE] silences sound. It does not extinguish fire.

\section{The Leap That Landed Wrong}

I once cast [LEAP] to cross a chasm. I had cast the spell a hundred times. This time, I was tired, and I omitted the [WITNESS] tag that I usually included. The Weave obliged. I landed exactly where I intended. Unfortunately, the space where I intended to land was already occupied by a boulder that had not been there when I cast. The Weave had not moved the boulder. It had simply \emph{not informed me}. I learned: a teleport is only as good as your knowledge of the destination. Scry before you leap.

\section{The Healing That Hurt}

A student in my care attempted [HEAL] on a wounded comrade. She rolled a miss. The Weave did not refuse to heal. It \emph{transferred} the wound. The comrade's injury vanished. The student's arm opened with the same gash. She learned: the Weave balances ledgers. Healing is not creation; it is \emph{movement}.

\section{The Veil That Veiled Too Much}

A confidence artist—not my student, thankfully—cast [VEIL] on his face during a negotiation. He wanted to make himself forgettable. The Weave obliged. His face blurred to the point that his own allies could not recognise him. The negotiation failed. He was stabbed by a guard who thought he was a different criminal. He survived. He learned: a veil should be measured. The Weave does not understand subtlety unless you specify.

\section{The Gate That Opened Sideways}

I have mentioned this failure elsewhere. A student cast [GATE] in an examination. The portal opened not to the adjacent hallway but to a room filled with mirrors. He saw versions of himself that had made different choices. He closed the portal. He withdrew from the University. He learned: the Weave's sense of destination is influenced by the caster's subconscious. If you are unsure of where you want to go, the Weave will guess. Its guesses are not kind.

\section{The Summoning That Succeeded}

A summoner—a colleague, not a student—called a spirit with [SUMMON]. The spirit arrived. It was exactly the spirit he had named. It was also, it turned out, the spirit of his own dead father, who had not spoken to him in life. The conversation lasted three hours. The summoner did not cast again for a year. He learned: a successful summoning is not necessarily a useful one. The Weave does not filter for comfort.

\section{What These Failures Teach}

\begin{itemize}
    \item \textbf{Know your tags.} Each tag names exactly one quality. Do not assume synonyms.
    \item \textbf{Know your destination.} [LEAP] and [GATE] require current knowledge, not hope.
    \item \textbf{Healing is transfer.} The Weave does not create; it reallocates.
    \item \textbf{Measure your veils.} Too much concealment conceals you from your allies.
    \item \textbf{The Weave reads your subconscious.} Cast with a clear mind or accept the consequences.
\end{itemize}

The Weave is not a machine. It is a mirror. Failures teach you what the mirror sees.

\textit{--- Lysandra}
