\chapter{Tags of the Weave}

This appendix presents the complete tag vocabulary in two forms. The first table groups tags by Difficulty Value—standard tags (DV +1) and dangerous tags (DV +2). The second table lists all tags alphabetically with their families and a brief description.

Dangerous tags are marked in \textcolor{red}{red}. The Weave charges a surcharge for these tags not because it disapproves, but because folding reality requires more ink.

\section{Tags by Difficulty Value}

\subsection{Standard Tags (DV +1)}

These tags add one to the spell's Difficulty. They are the everyday vocabulary of the Weave—common, reliable, and (mostly) safe.

\noindent
\begin{longtable}{p{0.2\textwidth} p{0.15\textwidth} p{0.55\textwidth}}
\caption{Standard Tags (DV +1)}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{What It Does}\\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{What It Does}\\
\midrule
\endhead
AREA & Kinetic & Expands a spell's effect to cover a zone. \\
BEAST & Life & Communicates with or influences an animal. \\
BIND & Kinetic & Restrains a target. \\
BLESS & Affliction & Grants a favourable tag—luck, favour, or ward- key. \\
BURNING & Elemental (Fire) & Ignites, heats, or smokes a target. \\
COMMAND & Mind & Compels a single short action. \\
COUNTER & Reaction & Interrupts a casting or ritual in its window. \\
CURSE & Affliction & Attaches a hostile tag or clock to a target. \\
DEFEND & Force & Creates a personal barrier. \\
DISPEL & Kinetic & Unravels an ongoing magical effect. \\
FEAR & Mind & Instils dread in a target. \\
FORCE & Kinetic & Pushes, pulls, or strikes with invisible pressure. \\
FREEZING & Elemental (Water) & Cools, chills, or encases in frost. \\
HEAL & Life & Restores Fatigue or mends minor wounds. \\
LIGHT & Utility & Creates illumination. \\
MEMORY & Mind & Touches the past—blurring, sharpening, or removing recollections. \\
PROTECT & Utility & Reduces or deflects next harm. \\
PURIFY & Life & Removes poison, disease, or corruption. \\
REVEAL & Utility & Unveils hidden, glamoured, or invisible things. \\
SCRY & Mind & Reveals a distant location or object. \\
SENSE & Utility & Detects the presence of a named tag or element. \\
SHADOW & Utility & Deepens darkness, hides edges, obscures. \\
SILENCE & Utility & Suppresses sound in a zone or on a target. \\
STONE & Elemental (Earth) & Hardens, weights, or shapes stone and earth. \\
STORM & Elemental (Air) & Calls lightning, thunder, or static charge. \\
STRENGTHEN & Life & Temporarily enhances a body. \\
STRIKE & Kinetic & Focuses a spell on a single target. \\
VEIL & Mind & Conceals a target from sight, memory, or attention. \\
WAKEN & Life & Rouses a sleeping or stunned creature. \\
WALL & Kinetic & Creates a barrier of force, element, or illusion. \\
WAVE & Elemental (Water) & Moves water or creates pressure. \\
WIND & Elemental (Air) & Moves air. \\
\hline
\end{longtable}

\subsection{Dangerous Tags (DV +2)}

\noindent
\begin{longtable}{p{0.2\textwidth} p{0.15\textwidth} p{0.55\textwidth}}
\caption{Dangerous Tags (DV +2)}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{What It Does}\\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{What It Does}\\
\midrule
\endhead
\textcolor{red}{ANIMATE} & Creation & Gives motion to an inanimate object. \\
\textcolor{red}{CREATE} & Creation & Manifests a mundane object from nothing. \\
\textcolor{red}{DOMINATE} & Mind & Imposes the caster's will on a sentient creature. \\
\textcolor{red}{FOLD} & Space & Bends space around an obstacle. \\
\textcolor{red}{GATE} & Space & Creates a two- way portal between two locations. \\
\textcolor{red}{GRAVITY} & Space & Alters local gravity. \\
\textcolor{red}{LEAP} & Space & Teleports the caster within Near range. \\
\textcolor{red}{REFLECT} & Reaction & Turns the next targeted effect back on its source. \\
\textcolor{red}{REWIND} & Time & Reverses a few heartbeats. \\
\textcolor{red}{STORE} & Utility & Banks successes in a vessel for later use. \\
\textcolor{red}{SUMMON} & Creation & Calls a spirit, elemental, or minor Outsider. \\
\textcolor{red}{TRANSMUTE} & Creation & Changes one substance into another. \\
\textcolor{red}{TRANSFORM} & Creation & Changes a creature's form. \\
\hline
\end{longtable}

\section{Tags Alphabetically}

For quick reference at the table.

\noindent
\begin{longtable}{p{0.2\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.4\textwidth}}
\caption{Tags Alphabetical}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{DV Mod} & \textbf{Effect}\\
\midrule
\endfirsthead
\multicolumn{4}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{DV Mod} & \textbf{Effect}\\
\midrule
\endhead
ANIMATE & Creation & +2 & Gives motion to an inanimate object. \\
AREA & Kinetic & +1 & Expands effect to a zone. \\
BEAST & Life & +1 & Communicates with or influences animals. \\
BIND & Kinetic & +1 & Restrains a target. \\
BLESS & Affliction & +1 & Grants a favourable tag. \\
BURNING & Elemental & +1 & Ignites, heats, or smokes. \\
COMMAND & Mind & +1 & Compels a single short action. \\
COUNTER & Reaction & +1 & Interrupts a casting or ritual. \\
CREATE & Creation & +2 & Manifests a mundane object from nothing. \\
CURSE & Affliction & +1 & Attaches a hostile tag or clock. \\
DEFEND & Force & +1 & Creates a personal barrier. \\
DISPEL & Kinetic & +1 & Unravels an ongoing magical effect. \\
DOMINATE & Mind & +2 & Imposes will on a sentient creature. \\
FEAR & Mind & +1 & Instils dread. \\
FOLD & Space & +2 & Bends space around an obstacle. \\
FORCE & Kinetic & +1 & Pushes, pulls, or strikes with pressure. \\
FREEZING & Elemental & +1 & Cools, chills, or encases in frost. \\
GATE & Space & +2 & Creates a two- way portal. \\
GRAVITY & Space & +2 & Alters local gravity. \\
HEAL & Life & +1 & Restores Fatigue, mends minor wounds. \\
LEAP & Space & +2 & Teleports the caster within Near range. \\
LIGHT & Utility & +1 & Creates illumination. \\
MEMORY & Mind & +1 & Touches the past—blurs, sharpens, removes. \\
PROTECT & Utility & +1 & Reduces or deflects next harm. \\
PURIFY & Life & +1 & Removes poison, disease, corruption. \\
REFLECT & Reaction & +2 & Turns next targeted effect back on source. \\
REVEAL & Utility & +1 & Unveils hidden, glamoured, or invisible things. \\
REWIND & Time & +2 & Reverses a few heartbeats. \\
SCRY & Mind & +1 & Reveals a distant location or object. \\
SENSE & Utility & +1 & Detects presence of a named tag or element. \\
SHADOW & Utility & +1 & Deepens darkness, hides edges. \\
SILENCE & Utility & +1 & Suppresses sound. \\
STONE & Elemental & +1 & Hardens, weights, or shapes stone and earth. \\
STORE & Utility & +2 & Banks successes in a vessel for later use. \\
STORM & Elemental & +1 & Calls lightning, thunder, static charge. \\
STRENGTHEN & Life & +1 & Temporarily enhances body. \\
STRIKE & Kinetic & +1 & Focuses spell on a single target. \\
SUMMON & Creation & +2 & Calls a spirit, elemental, or Outsider. \\
TRANSMUTE & Creation & +2 & Changes one substance into another. \\
TRANSFORM & Creation & +2 & Changes a creature's form. \\
VEIL & Mind & +1 & Conceals from sight, memory, attention. \\
WAKEN & Life & +1 & Rouses sleeping or stunned creature. \\
WALL & Kinetic & +1 & Creates a barrier. \\
WAVE & Elemental & +1 & Moves water or creates pressure. \\
WIND & Elemental & +1 & Moves air. \\
\hline
\caption*{Dangerous tags (DV +2) are marked in \textcolor{red}{red} in the main table. The Weave charges a surcharge for these effects. Pay it or do not cast.}
\end{longtable}

\section{On Tag Combinations}

Some tags work well together. [BURNING] and [AREA] produce a conflagration. [SILENCE] and [AREA] produce a zone of quiet. [LEAP] and [STRIKE] produce a teleporting attack—if you can manage the DV.

Some tags conflict. [HEAL] and [CURSE] cannot occupy the same spell; the Weave will choose one and ignore the other. [VEIL] and [REVEAL] cancel each other; the Weave will laugh and charge you for both.

Experiment in a safe space. Take notes. The Weave will not punish honest experimentation. It will punish carelessness.

\textit{--- Lysandra}
