\chapter{The Menagerie's Mortgage}
łabel{chap:menagerie−mortgage}

\section*{Adventure Overview}
\textit{A surreal Fate's Edge parody adventure in a pocket−dimension safari gone horribly off−budget, morally questionable, and faintly sticky.}

Maelstraeus, the Infernal Bargainer, hires the party to serve a legally−binding summons to \textbf{Zara the Magnificent}, owner of the \textbf{Celestial Menagerie} −−− a chain of pocket−dimensional enclosures filled with ``educational'' exhibits that mostly educate visitors about trauma recovery.

Zara has missed multiple payments on her infernal mortgage and fled into her own park. To complete the contract, the summons must be served \textit{directly} to Zara before she escapes the Menagerie entirely.

\section*{Core Hook}
\begin{quote}
``The contract says we must serve her personally. It does not specify dimension, altitude, or whether she has grown extra teeth.''
\end{quote}

\section{The Contract}
The summons is a magical document with unfortunate opinions:
\begin{itemize}
  ℹtem Burns anyone attempting to tear it.
  ℹtem Screams loudly when ignored.
  ℹtem Glows brighter when closer to Zara.
  ℹtem Fine print is literally microscopic (Arcana DV~3 to decipher).
\end{itemize}

\textbf{Failure:} If not delivered within 7 days, the party owes Maelstraeus a favor.  
\textbf{Success:} Zara owes \textit{them} a favor. She is \textit{very} creative with repayment.

\section{Adventure Structure}
The Celestial Menagerie consists of 12 loosely regulated enclosures. Each is a self−contained scene with hazards, social encounters, or problems that are technically ``educational,'' depending on one's definition of ``education.''

\bigskip

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\section{Enclosures}

% −−−−−−−−−− 1 −−−−−−−−−−
\subsection{Enclosure 1: The Candle−Lizard Warrens}
\textbf{Theme:} Trap−building lizard−folk with explosive candle helmets.

\textbf{Challenge:} A maze of low tunnels, trick floors, and unhelpful residents.

\begin{description}
  ℹtem[Initial Position:] Controlled
  ℹtem[Position Shift:] Controlled $→$ Desperate (bribing or befriending a traitorous lizard)
\end{description}

\textbf{Story Beat (SB) Spends:}
\begin{itemize}
  ℹtem 1 SB: Party−triggered springboard launches someone backwards.
  ℹtem 2 SB: Idol of the Ancient Fire−Chicken shrieks and panics nearby creatures.
  ℹtem 3 SB: Tunnel collapse, Environmental Timer  +1.
\end{itemize}

% −−−−−−−−−− 2 −−−−−−−−−−
\subsection{Enclosure 2: The Hungry Furniture Exhibition}
\textbf{Theme:} Furniture that hungers for attention and also flesh.

\textbf{Challenge:} Identifying real exhibits from predators in disguise.

\begin{description}
  ℹtem[Initial Position:] Desperate
  ℹtem[Position Shift:] Desperate $→$ Controlled (spotting tells in the upholstery)
\end{description}

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: ``Treasure chest'' bites.
  ℹtem 2 SB: Informational plaques are also carnivorous.
  ℹtem 3 SB: Curator−Cabinet awakens and lectures the party about ``proper museum etiquette.''
\end{itemize}

% −−−−−−−−−− 3 −−−−−−−−−−
\subsection{Enclosure 3: The Rust−Bugs' Picnic Grounds}
\textbf{Theme:} Enthusiastic metal−eaters.

\textbf{Challenge:} Protecting equipment while maintaining dignity.

\begin{description}
  ℹtem[Initial Position:] Desperate
  ℹtem[Position Shift:] Desperate $→$ Controlled (using wood, bone, or stone tools)
\end{description}

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Weapon gains the ``Neglected'' condition.
  ℹtem 2 SB: Armor becomes ``Compromised.''
  ℹtem 3 SB: Magical metal loses enchantment layers.
\end{itemize}

% −−−−−−−−−− 4 −−−−−−−−−−
\subsection{Enclosure 4: The Gelatin Menagerie}
\textbf{Theme:} Digestive oozes and transparent hazards.

\textbf{Challenge:} Avoiding invisible cubes and retrieving swallowed equipment.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Item swallowed (Tinker DV 3 to recover).
  ℹtem 2 SB: Slimes merge into a Large Cube.
  ℹtem 3 SB: Acid pool spreads (Harm~1 area).
\end{itemize}

% −−−−−−−−−− 5 −−−−−−−−−−
\subsection{Enclosure 5: The Goblin Circus of Mild Regret}
\textbf{Theme:} Chaotic performers with a strong union and no hazard pay.

\textbf{Challenge:} Pranks, slapstick, and forced audience participation.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Cream−pie to the face (Position −−1).
  ℹtem 2 SB: ``Elephant'' (three goblins in a costume) stampedes.
  ℹtem 3 SB: Ringmaster demands a heroic performance or duel.
\end{itemize}

% −−−−−−−−−− 6 −−−−−−−−−−
\subsection{Enclosure 6: Owlbear−Adjacent Preservation Zone}
\textbf{Theme:} Creatures that are definitely not owls and definitely not bears, legally speaking.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: ``Affectionate'' grapple (Harm~1).
  ℹtem 2 SB: Cubs summon protective parent.
  ℹtem 3 SB: Containment fence breaks.
\end{itemize}

% −−−−−−−−−− 7 −−−−−−−−−−
\subsection{Enclosure 7: Corroded Metal Dragon Roost}
\textbf{Theme:} Dragons with mortgage−based emotional distress.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Corrosive breath affects metal gear.
  ℹtem 2 SB: Dragon spirals into existential crisis.
  ℹtem 3 SB: Rust−bugs arrive for a feast.
\end{itemize}

% −−−−−−−−−− 8 −−−−−−−−−−
\subsection{Enclosure 8: Psychic Jellyfish Sanctuary}
\textbf{Theme:} Telepathic puffballs who deeply regret overhearing thoughts.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Puffballs broadcast embarrassing secrets.
  ℹtem 2 SB: Psychic feedback (Position −−1).
  ℹtem 3 SB: A tentacled brain−horror breaches containment.
\end{itemize}

% −−−−−−−−−− 9 −−−−−−−−−−
\subsection{Enclosure 9: The Living Armor Parade}
\textbf{Theme:} Security suits that gained collective bargaining rights.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Defeated armor reassembles.
  ℹtem 2 SB: Exhibit security lockdown.
  ℹtem 3 SB: Ancient Elite Armor demands paperwork.
\end{itemize}

% −−−−−−−−−− 10 −−−−−−−−−−
\subsection{Enclosure 10: Phase−Cat Habitat}
\textbf{Theme:} Illusionary predators with a superiority complex.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Illusionary wounds (Position −−1).
  ℹtem 2 SB: Labyrinth shifts.
  ℹtem 3 SB: Real Phase−Cat strikes.
\end{itemize}

% −−−−−−−−−− 11 −−−−−−−−−−
\subsection{Enclosure 11: Forbidden Psionics Research Wing}
\textbf{Theme:} Too many brains, not enough ethics board.

\textbf{SB Spends:}
\begin{itemize}
  ℹtem 1 SB: Psychic migraine (Harm~1 mental).
  ℹtem 2 SB: Mind−control attempt (DV~3 resist).
  ℹtem 3 SB: Failed experiment mutates and wants hugs.
\end{itemize}

% −−−−−−−−−− 12 −−−−−−−−−−
\subsection{Enclosure 12: Zara's Office}
\textbf{Theme:} The final confrontation with a legally−minded illusionist.

\textbf{Endings:}
\begin{itemize}
  ℹtem \textbf{Serve the Summons:} Success; Zara owes a favor.
  ℹtem \textbf{Take Maelstraeus's Deal:} Immediate reward, new rival.
  ℹtem \textbf{Rewrite Contract:} Benefit both sides (high DV).
  ℹtem \textbf{Violence:} Zara escapes; no reward; plot hook unlocked.
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\section{Attraction Tickets}
Players earn \textbf{Attraction Tickets} via creative problem−solving, social play, or making the GM laugh.

\textbf{Tickets May Be Spent To:}
\begin{itemize}
  ℹtem Skip one enclosure entirely.
  ℹtem Improve Position by one step.
  ℹtem Gain a clue from Maelstraeus.
  ℹtem Escape a catastrophic situation.
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\section{Why This Works}
\begin{enumerate}
  ℹtem Classic creature archetypes reimagined as parody−friendly encounters.
  ℹtem Multiple paths: social, stealth, violent, and absurd.
  ℹtem Ticket economy encourages creativity and agency.
  ℹtem Modular structure works for one−shots or full arc.
  ℹtem Humor integrated mechanically, not just narratively.
\end{enumerate}

\section{Souvenir Shops & Generators}
łabel{sec:souvenir−shops}

\subsection*{Using Souvenir Shops}
Souvenir shops appear near enclosure exits, tram stops, or panic doors. They provide comic relief, light gear, edible mistakes, and surprisingly binding receipts.

\paragraph{Core Loops.}
\begin{itemize}
  ℹtem \textbf{Attraction Tickets:} May be used as currency (see Table~\ref{tab:ticket−exchange}).
  ℹtem \textbf{Haggling:} \emph{Sway} or \emph{Trade} (DV 2−−3) to modify price or perk.
  ℹtem \textbf{Receipt Law:} All receipts are extradimensionally enforceable. Fine print always matters.
  ℹtem \textbf{Safety:} Shops \emph{seem} safe. SB spends can invalidate that assumption.
\end{itemize}

\paragraph{SB Hooks (GM).}
On SB spends in a shop scene, choose or roll:
\begin{itemize}
  ℹtem Shelf domino cascade (Position −−1)  |  Counterfeit tickets discovered (Lose 1 Ticket)
  ℹtem Clerk calls security armor (see Enclosure 9)  |  A ``demo curse'' triggers (Harm 1 or Tag)
  ℹtem Warranty clause teleports item back to store at worst time
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\subsection{Ticket Exchange & Pricing}
łabel{subsec:ticket−exchange}

\begin{table}[h]
¢ering
∩tion{Attraction Ticket Exchange (per shopper per shop)}
łabel{tab:ticket−exchange}
\begin{tabular}{@{}ll@{}}
⊤rule
\textbf{Spend} & \textbf{Effect} \\
∣rule
1 Ticket & Minor trinket or snack; +1d to next social roll in this shop \\
2 Tickets & Useful consumable (one−scene tag), or reroll one failure on haggling \\
3 Tickets & Curio capsule (see Table~\ref{tab:curio−gacha}) or shop−only boon \\
5 Tickets & Pick any one result from \emph{Trinkets} or \emph{Mildly Cursed Merch} \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{Baseline Prices}
łabel{tab:baseline−prices}
\begin{tabular}{@{}ll@{}}
⊤rule
\textbf{Item Tier} & \textbf{Typical Cost} \\
∣rule
Snack / Trinket & 1−−2 coin \emph{or} 1 Ticket \\
Curio Capsule & 3 coin \emph{or} 2 Tickets \\
Mildly Cursed Merch & 6−−9 coin \emph{or} 3−−4 Tickets \\
Warranty Add−on & +1−−3 coin (see Table~\ref{tab:warranty}) \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\subsection{Shopframe Generators (Roll What You Need)}

\begin{table}[h]
¢ering
∩tion{d8 Shop Theme}
łabel{tab:shop−theme}
\begin{tabular}{@{}cl@{}}
⊤rule
d8 & Theme \\
∣rule
1 & Glowsticks & Bioluminescence Boutique \\
2 & Comfy Monsters Plush Emporium \\
3 & Legalese & Stationery Kiosk \\
4 & Pocket Provisions (snacks, tonics, regrets) \\
5 & ``Authentic'' Replicas & Mini Models \\
6 & Curio Capsule & Gacha Corner \\
7 & Lost & Found & For Sale \\
8 & Clearance Curses Warehouse \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{d12 Clerk Persona}
łabel{tab:clerk−persona}
\begin{tabular}{@{}cl@{}}
⊤rule
d12 & Persona \\
∣rule
1 & Exhausted grad of Infernal Business School, knows loopholes \\
2 & Hyper−sincere mimic wearing a name−tag (definitely a clerk) \\
3 & Retired security armor with customer−service voice \\
4 & Telepathic puffball who overshares discounts \\
5 & Candle−lizard intern; flammable enthusiasm \\
6 & Dragon accountant doing community service \\
7 & Goblin ringmaster moonlighting for store credit \\
8 & Gelatin cube with tray; communicates via squeaky sign board \\
9 & Ex−thrall memorizing policy manual verbatim \\
10 & Phase−cat in a vest; appears before questions are asked \\
11 & ``Assistant manager''−−−a sentient receipt stapler \\
12 & Zara's hologram; can only sell warranty plans \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{d12 Shop Quirk}
łabel{tab:shop−quirk}
\begin{tabular}{@{}cl@{}}
⊤rule
d12 & Quirk \\
∣rule
1 & Prices change with moon phase (or nearest glowing sign) \\
2 & Every tenth customer becomes ``Employee of the Minute'' (10% off) \\
3 & Door locks during sales pitches \\
4 & Samples are mildly sentient and ask for reviews \\
5 & Shelves re−arrange if you look away \\
6 & Receipt prints prophecies in 6pt font \\
7 & Floor is a slow conveyor belt toward the register \\
8 & Anti−theft runes occasionally applaud \\
9 & Background music is legally binding \\
10 & Fittings mirror shows possible futures (Position +1 once) \\
11 & Loyalty card requires a blood \emph{or} jam thumbprint \\
12 & Returns desk is in a pocket closet two steps to the left of reality \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\subsection{Merch Tables}

\begin{table}[h]
¢ering
∩tion{d20 Trinkets & Pocket Curios}
łabel{tab:trinkets}
\begin{tabular}{@{}cl@{}}
⊤rule
d20 & Item (Effect) \\
∣rule
1 & Glow−mote vial (one scene: create Light tag in Close range) \\
2 & Foldable paper familiar (once: +1d to Scout, then soggy) \\
3 & Pocket map that updates only after you get lost (reduce Time cost by 1) \\
4 & Miniature cube scrubber (clean any item; removes one grime−related penalty) \\
5 & Sticker of authority (once: +1 effect on bureaucrats) \\
6 & Safety whistle audible to security armor (summon 1 suit; Position +1 or −−1) \\
7 & Ribbon of the Ringmaster (once: reroll one failure in Perform) \\
8 & Dragon−IOU punch card (3 punches = 1 trivial favor from a bored drake) \\
9 & Phase−cat bell (once: negate an illusion penalty) \\
10 & Puffball postcard (read surface thought of a willing target) \\
11 & "I Survived the Menagerie" pin (once: resist Harm 1 from embarrassment) \\
12 & Pocket legal pad (once: introduce a minor clause; +1d to Sway) \\
13 & Rust−bug treat bag (once: pacify rust−bugs for one scene) \\
14 & Candle−lizard sparkler (create Distract tag for one roll) \\
15 & Counterfeit badge (on success: +1 Position; on failure: security called) \\
16 & Telescoping reacher (tag: Long Handle; +1 effect on hard−to−reach grabs) \\
17 & Collapsible ladder (2−segment climb shortcut) \\
18 & Souvenir poncho (ignore one splatter/acid splash penalty) \\
19 & Mini−wand of pointing (once: ask GM "where is the nearest safe exit?") \\
20 & Pocket translator (once: speak one enclosure's critter pidgin) \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{d12 Snack Stall Specials}
łabel{tab:snacks}
\begin{tabular}{@{}cl@{}}
⊤rule
d12 & Snack (Effect) \\
∣rule
1 & Infernal kettle−corn (focus: +1d next Insight) \\
2 & Jellylight parfait (glow: you emit Dim Light for a scene) \\
3 & Rust−proof jerky (gear: ignore one rust event this scene) \\
4 & Puffball cotton candy (calm: reduce Stress by 1, then sugar crash Position −−1) \\
5 & Phase−cat macarons (once: ignore one illusion penalty) \\
6 & Goblin churro (bravado: +1 effect on Perform) \\
7 & Dragon−breath mints (minty: speak without stammer to authority; +1d Sway) \\
8 & Cube jello shots (resist one splash−based Harm 1) \\
9 & Candle−lizard hot sauce (apply Burn tag to your next thrown snack?) \\
10 & "Authentic" nut−free trail mix (heal 1 Fatigue if seated 10 minutes) \\
11 & Armor−oil espresso (perk: first jammed mechanism is suddenly smooth) \\
12 & Warranty wafer (file a \emph{minor} claim to undo a trivial mishap) \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{d20 Curio Capsule & Gacha Corner}
łabel{tab:curio−gacha}
\begin{tabular}{@{}cl@{}}
⊤rule
d20 & Capsule Prize (Effect) \\
∣rule
1 & Mini fog machine bead (one scene: create Cover tag) \\
2 & Pocket rope snake (self−knotting; +1 effect on Bind/Tie) \\
3 & Micro grapnel coin (once: traverse a Short gap safely) \\
4 & Complaint token (trade to clerk: +1 tier item upgrade) \\
5 & Refund chit (cancel one SB spend in a shop) \\
6 & "Manager Override" badge (once: treat DV as 1 for a purchase) \\
7 & Safety rune decal (once: reduce incoming Harm 2 to Harm 1) \\
8 & Contract staple of holding (attach any paper to any surface, forever) \\
9 & Dramatic cape clip (once: +1d to Impress/Intimidate) \\
10 & Accurate souvenir map (ignore one maze reposition) \\
11 & Snack voucher (free roll on Table~\ref{tab:snacks}) \\
12 & Mini puzzle box (spend 10 minutes: clear 1 Stress) \\
13 & Pocket siren (disperse small crowd or draw help, GM's choice) \\
14 & Anti−slip charms (ignore first banana/floor hazard this scene) \\
15 & Echo pebble (repeat last heard phrase; good for passwords) \\
16 & Deedletag (label an item; find it across enclosures once) \\
17 & Emergency tram whistle (summon slow transport; buys time) \\
18 & Fine−print magnifier (read any tiny clause without DV) \\
19 & VIP wristband (once: skip a queue or enclosure gate) \\
20 & Golden capsule (choose any result on this table) \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{d12 Mildly Cursed Merchandise}
łabel{tab:cursed}
\begin{tabular}{@{}cl@{}}
⊤rule
d12 & Item (Benefit & Quirk) \\
∣rule
1 & Sword of Customer Loyalty (+1 effect in this shop; cannot \emph{leave} the shop willingly) \\
2 & T−Shirt of Hubris (once/day: +1d any roll; on failure, shirt yells) \\
3 & Mug of Endless Lukewarm (water forever; never hot or cold) \\
4 & Hat of Confident Directions (+1d to Navigate; always off by one hallway) \\
5 & Cloak of Legal Shade (+1d vs. contracts; attracts mall cops) \\
6 & Sandals of Soft Steps (silent walk; squeak loudly on lies) \\
7 & Bag of Holding(ish) (tiny extradimensional pocket; occasionally returns confetti) \\
8 & Lens of Flaw−Finding (+1d to Inspect; bearer becomes tediously picky) \\
9 & Plush of Emotional Support (clear 1 Stress when hugged; demands weekly check−ins) \\
10 & Umbrella of Dry Irony (rain misses you; drinks spill toward you) \\
11 & Key of Most Doors (opens 1 mundane lock/day; loudly announces itself) \\
12 & Warranty Badge of Doom (free returns; receipts chase you across planes) \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\subsection{Policies, Warranties, and Fine Print}

\begin{table}[h]
¢ering
∩tion{d12 Warranty Add−Ons & Clauses}
łabel{tab:warranty}
\begin{tabular}{@{}cl@{}}
⊤rule
d12 & Clause (Mechanic) \\
∣rule
1 & 24−Hour Satisfaction: once, undo last purchase consequence \\
2 & Accidental Oozing: immune to slime stains for one scene \\
3 & Teleport−to−Register: item returns to shop if broken (at worst time) \\
4 & Extended Fine Print: treat one contract DV as 1 \\
5 & Manager's Special: one reroll on Tables \ref{tab:trinkets}−−\ref{tab:cursed} \\
6 & Loyalty Upgrade: after three purchases, free roll on Gacha \\
7 & Anti−Theft Rune: item screams when stolen (by anyone) \\
8 & Cross−Plane Coverage: valid in any shop sharing brand sigil \\
9 & Goblin Repair Plan: fix one non−magic item between scenes \\
10 & Consumables Reset: refresh one consumed snack or curio \\
11 & Safety Recall: once, convert Harm 2 (shop−related) to a funny story \\
12 & Legal Counsel On Call: once, +1d vs. Zara's office antics \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{d10 Returns & Exchanges Outcome}
łabel{tab:returns}
\begin{tabular}{@{}cl@{}}
⊤rule
d10 & Outcome \\
∣rule
1 & Full refund, unexpected store credit equal to 1 Ticket \\
2 & Exchange approved; clerk adds free snack sample \\
3 & Partial refund in coins that try to crawl away \\
4 & Manager signature required (mini social scene; Position +1 if flattered) \\
5 & Approved but receipt now follows you, whispering advice \\
6 & Denied on technicality; fine print reveals side quest for approval \\
7 & Restocking fee: lose 1 Ticket \\
8 & Item teleports back to shelf mid−argument \\
9 & Store audit: security armor "assists" the process \\
10 & Apology parade: gain 1 Ticket, but become shop's VIP for 24h \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\subsection{Haggling, Sales, and Events}

\begin{table}[h]
¢ering
∩tion{d10 Price Scheme & Haggling Hooks}
łabel{tab:price−scheme}
\begin{tabular}{@{}cl@{}}
⊤rule
d10 & Scheme (Hook) \\
∣rule
1 & Color−tag sale (predict next tag; DV 2 to time purchase) \\
2 & Mystery discount wheel (spin for −−10% to −−▒) \\
3 & Bundle pricing (buy 2, get +1d on one) \\
4 & Ticket match (spend Tickets = coin discount) \\
5 & Clerk's discretion (befriend them for −−░) \\
6 & Timed flash sale (two rounds only; Position −−1 from rush) \\
7 & Loyalty ladder (cumulative purchases unlock better tiers) \\
8 & Receipt riddle (solve for −−30%) \\
9 & Trade−in accepted (swap one trinket for upgrade) \\
10 & Charity jar (donate 1 coin: cancel one shop SB spend) \\
⊥tomrule
\end{tabular}
\end{table}

\begin{table}[h]
¢ering
∩tion{d8 Flash Events (While You Browse)}
łabel{tab:flash−events}
\begin{tabular}{@{}cl@{}}
⊤rule
d8 & Event \\
∣rule
1 & Shelf avalanche (Group Move DV 2 to avoid slapstick) \\
2 & Surprise mascot conga line (join for +1d Perform or be blocked) \\
3 & Power flicker; prices randomize (reroll Table~\ref{tab:price−scheme}) \\
4 & VIP tour barges through (Sway DV 2 to glean gossip about Zara) \\
5 & Demonstration of cursed item goes awry (Harm 1 unless duck) \\
6 & Intercom summons "new hire" (one PC becomes clerk for 1 scene) \\
7 & Secret shopper evaluates you (Impress for hidden perk) \\
8 & Fire−safety drill led by candle−lizard (Position +1 if you help) \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\subsection{Quick Mechanical Tags (Reference)}
\begin{itemize}
  ℹtem \textbf{Light:} You can see; reduce darkness penalties in Close range.
  ℹtem \textbf{Cover:} Incoming ranged harm reduced by one step once.
  ℹtem \textbf{Distract:} Next enemy roll suffers reduced effect.
  ℹtem \textbf{Long Handle:} Interact safely at Close from just outside danger.
\end{itemize}

\paragraph{GM Integrations.}
\begin{itemize}
  ℹtem \textbf{Zara's Escalation Timer :} +1 on major shop incidents (alarms, audits).
  ℹtem \textbf{Exhibit Maintenance:} −−1 when players buy repair kits or volunteer aid.
  ℹtem \textbf{Favor (Maelstraeus):} Award +1 pip if players spot/exploit funny contractual clauses.
\end{itemize}

\section{Zara's Contract of Infernal Mortgage and Adjacent Obligations}
łabel{sec:zara−contract}

\begin{tcolorbox}[title={¢ering \textbf{Celestial Menagerie −− Infernal Mortgage Contract}},colback=white,colframe=black!60]
\begin{center}
{Łarge The Celestial Menagerie \\
Mortgage & Summons−Delivery Agreement}
\end{center}

\noindent
\textbf{By Authority of the Infernal Bargaining Office, Fourth Subdivision, Under−Market Division of Contracts and Culinary Arts}

\vspace{0.2cm}

\noindent
Let it be publicly, cosmically, and awkwardly known that:

\begin{enumerate}
  ℹtem \textbf{Zara the Magnificent}, legal owner and sole visionary of The Celestial Menagerie (``the Debtor''), did on the Twenty−Third Cycle of the Glowing Chicken Moon, Year of the Infernal Spreadsheet, sign a Mortgage of Infernal Standing.
  ℹtem Said Mortgage provided funds, labor, pocket−dimension stitching, signage, pamphlets, and optional snack kiosks.
  ℹtem Payments were due on each market day, excluding Feastdays, Half−Feastdays, Minor Feastday Eves, and the Day of Spontaneous Picnics.
\end{enumerate}

\noindent
\textbf{Whereas} the Debtor has failed to render timely payment for \textit{three} consecutive cycles, the following actions are hereby authorized and encouraged:

\begin{enumerate}[label=\Alph*.]
  ℹtem Delivery of a Summons directly to the Debtor, physically, verbally, or magically;
  ℹtem Delivery may occur in any enclosure, sub−enclosure, office, attic, broom closet, or makeshift hot−dog stand;
  ℹtem Illusions, clones, impersonators, or holograms of the Debtor \textit{do not} count as valid recipients unless the Debtor has legally declared them ``emotionally equivalent'' representatives.
\end{enumerate}

\subsection*{Accepted Methods of Service}
\begin{itemize}
  ℹtem Handing the Summons to the Debtor's actual hands, talons, pseudopods, or other manipulators;
  ℹtem Telepathic transfer, \emph{if} the Debtor acknowledges receipt in complete sentences;
  ℹtem Firing the Summons point−blank into the Debtor's office wall via crossbow, \textit{if} the Debtor signs the impact crater.
\end{itemize}

\subsection*{Forbidden Methods of Service}
\begin{itemize}
  ℹtem Leaving the Summons on a desk, even if the Debtor is ``totally coming right back''; 
  ℹtem Handing it to a passing goblin and hoping for the best;
  ℹtem Stapling to the Debtor while invisible, flying, or disguised as interpretive dancers;
  ℹtem Service by \emph{vibe}.
\end{itemize}

\subsection*{Consequences of Valid Delivery}
\begin{enumerate}
  ℹtem The Debt shall become \textbf{immediately due}, with applicable late fees, referral fees, and a small surcharge for ``audience discomfort.''
  ℹtem The Debtor must either:
  \begin{itemize}
    ℹtem Repay the Mortgage in full;
    ℹtem Enter agreed−upon negotiations with the Infernal Bargainer;
    ℹtem Offer a Favor of Equal or Greater Value to the Deliverers (``the Bearers of Paper'').
  \end{itemize}
  ℹtem The Bearers of Paper shall have first right of refusal on any Favor of absurd or sticky nature.
\end{enumerate}

\subsection*{Consequences of Failure to Deliver}
\begin{itemize}
  ℹtem Upon failure to serve within seven solar days, the Bearers of Paper agree to render a Favor of Equal Weight.
  ℹtem The Infernal Bargainer shall, at their discretion, define ``Equal Weight'' as:
  \begin{enumerate}[label=\roman*.]
    ℹtem A simple errand involving bureaucratic paperwork,
    ℹtem A complicated errand involving bureaucratic paperwork,
    ℹtem Escort duty in a location with suspiciously high carnivore density,
    ℹtem Participation in a mandatory safety presentation lasting \textbf{four hours}.
  \end{enumerate}
\end{itemize}

\subsection*{Permissible Collateral Damage}
\begin{itemize}
  ℹtem Breakage of non−souvenir furniture;
  ℹtem Trampling of ornamental shrubs;
  ℹtem Damage to carnival stalls valued under 10 coin;
  ℹtem Emotional trauma inflicted on unpaid interns.
\end{itemize}

\subsection*{Prohibited Collateral Damage}
\begin{itemize}
  ℹtem Destruction of the Gift Shop's Profit Margin;
  ℹtem Harm to the Telepathic Puffballs (they unionized);
  ℹtem Collapsing more than one (1) pocket dimension at a time.
\end{itemize}

\vspace{0.2cm}

\begin{center}
\rule{0.8łinewidth}{0.4pt}\\
\textbf{Zara the Magnificent} (Debtor) ℎfill \textbf{Maelstraeus} (Infernal Bargainer)
\end{center}

\vspace{0.2cm}

\noindent
\textbf{Fine Print (Illegible to Naked Eye):}

\scriptsize
\begin{quote}
By reading this sentence, the reader agrees that any disputes shall be resolved by interpretive dance arbitration. All participants must provide their own tap shoes. This contract may scream if ignored. If accidentally eaten, contract grows back within 1d4 hours. Teleportation of contract without consent is punishable by irony. Warranties void in the presence of enchanted dairy.
\end{quote}

\end{tcolorbox}

\section{Location Generator (Celestial Menagerie)}
łabel{sec:location−generator}

\noindent
Use this generator when players wander off the main path, break containment fences, or deliberately explore ``staff−only'' areas. Each location has:
\begin{itemize}
  ℹtem A \textbf{Feature} (what is here?)
  ℹtem A \textbf{Hazard} (what could go wrong?)
  ℹtem A \textbf{Hook} (why does the location matter?)
  ℹtem A \textbf{Clock or Tag} to tie mechanically into Fate's Edge
\end{itemize}

\subsection*{Quick Roll Procedure}
\begin{enumerate}
  ℹtem Roll on Table~\ref{tab:location−type} for the broad zone.
  ℹtem Roll on Table~\ref{tab:location−feature} for the primary feature.
  ℹtem Roll on Table~\ref{tab:location−hazard} for the active problem.
  ℹtem Optionally roll on Table~\ref{tab:location−hook} for why players care.
\end{enumerate}

\vspace{0.5cm}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\begin{table}[h]
¢ering
∩tion{d12 Location Type}
łabel{tab:location−type}
\begin{tabular}{@{}cl@{}}
⊤rule
d12 & Zone Type \\
∣rule
1 & Overgrown staff trail lined with broken signage \\
2 & Maintenance catwalk above dangerous enclosure \\
3 & Cracked pocket−dimension seam leaking weird light \\
4 & Abandoned souvenir cart (half−eaten snacks) \\
5 & Temporary holding pens for escaped creatures \\
6 & Closed exhibit with tarp and ``Grand Opening Soon!'' sign \\
7 & Tram station with blinking lights and no tram \\
8 & Employee break room (sentient vending machine) \\
9 & Old storage warehouse with mislabeled crates \\
10 & Under−park utility tunnels with humming arcane pipes \\
11 & Lost−and−found heap organized by sentient coat rack \\
12 & VIP skybridge with shattered viewing crystals \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\begin{table}[h]
¢ering
∩tion{d20 Feature}
łabel{tab:location−feature}
\begin{tabular}{@{}cl@{}}
⊤rule
d20 & Feature \\
∣rule
1 & A shimmering portal window showing another enclosure \\
2 & Panic buttons that trigger the wrong alarms \\
3 & Cage full of glittering dart frogs (adorable, mildly poisonous) \\
4 & Trembling floor tiles (pressure triggers moving walkways) \\
5 & Oversized plush mascots that occasionally blink \\
6 & Incomplete holographic guide stuck in a feedback loop \\
7 & A fountain that produces flavored water (random effects) \\
8 & Creature feeding chute with abandoned snack pellets \\
9 & Pile of confiscated toy weapons (some real) \\
10 & Half−working cleaning drones dragging slime buckets \\
11 & A wall−length map with stickers marking ``incidents'' \\
12 & A collapsed viewing dome swarming with tiny glowing beetles \\
13 & A ``Do Not Touch'' exhibit rope placed around thin air \\
14 & A crashed delivery tram wedged sideways in a wall \\
15 & A ticket counter run by enchanted paper cut−outs \\
16 & Unrefrigerated cooler full of psychic jelly desserts \\
17 & Disguise booth with unstable illusion projectors \\
18 & A deflated bounce−castle habitat (dangerously stretchy) \\
19 & Giant gift−wrapped box rocking ominously from within \\
20 & A decorative wishing well that echoes back answers \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\begin{table}[h]
¢ering
∩tion{d20 Hazard}
łabel{tab:location−hazard}
\begin{tabular}{@{}cl@{}}
⊤rule
d20 & Hazard \\
∣rule
1 & Alarm runes activate (Position −−1 or security armor arrives) \\
2 & Floor gives way into slime−filled crawlspace \\
3 & Panicked creature stampedes through \\
4 & Illusory duplicates of the party appear and mock them \\
5 & Arcane sprinklers spray adhesive foam \\
6 & Staff−only drone demands identification and proof of snack purchase \\
7 & Pocket−dimension seam widens (Environmental Timer  +1) \\
8 & Gravity fluctuates (leap or float unpredictably) \\
9 & Invisible cleaner−cubes vacuum unattended items \\
10 & Support beams groan; structural collapse pending \\
11 & Animated plush tries to ``escort'' party to gift shop \\
12 & Frog chorus unleashes psychic stun (Harm 1 mental) \\
13 & Crystal shards detonate with harmless but blinding light \\
14 & Tram suddenly arrives at unsafe speed \\
15 & Feeding chute dispenses carnivorous snacks \\
16 & Slime leak creates slippery surface (fall risk) \\
17 & Anti−theft field teleports gear away briefly \\
18 & Dimensional static causes hallucinated voices \\
19 & Swarm of tiny mechanical parrots records everything \\
20 & Meanwhile, something inside that big gift−wrapped box wakes up \\
⊥tomrule
\end{tabular}
\end{table}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\begin{table}[h]
¢ering
∩tion{d12 Player−Relevant Hook}
łabel{tab:location−hook}
\begin{tabular}{@{}cl@{}}
⊤rule
d12 & Why This Matters \\
∣rule
1 & Clue to Zara's location hidden on a sign or recording \\
2 & Ticket cache hidden under counter (1d3 Tickets) \\
3 & Worker logbook reveals scandalous mismanagement \\
4 & Maintenance switch shortens path through next enclosure \\
5 & Creature egg, lost cub, or baby ooze needs help \\
6 & Chance to shut down alarm or slow Zara's escape timer \\
7 & Survivor or employee offers guidance if rescued \\
8 & Evidence of infernal contract loophole \\
9 & Shortcut portal to another enclosure \\
10 & Prized relic or shiny mascot−head worth coin or leverage \\
11 & Mirror that briefly shows Zara's future move \\
12 & Dimensional vent leading to secret mini−exhibit \\
⊥tomrule
\end{tabular}
\end{table}

\vspace{0.25cm}

\subsection*{Mechanical Notes (GM)}
\begin{itemize}
  ℹtem \textbf{Zara's Escalation Timer :} +1 when hazards trigger alarms, damage infrastructure, or release creatures.
  ℹtem \textbf{Exhibit Maintenance Timer :} −−1 if players repair, safely clear creatures, or deactivate hazards.
  ℹtem \textbf{Position:} Many locations start at \emph{Controlled} or \emph{Desperate} due to environmental chaos.
  ℹtem \textbf{Reward Tags:} \emph{Shortcut}, \emph{Cover}, \emph{Map Update}, \emph{Temporary Ally}.
\end{itemize}