\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{framed}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}

% Page geometry
\geometry{
  a4paper,
  margin=1.2in,
  top=1in,
  bottom=1in
}

% Colors
\definecolor{gold}{RGB}{201, 169, 76}
\definecolor{darkbg}{RGB}{26, 26, 26}
\definecolor{openbg}{RGB}{26, 58, 26}
\definecolor{closedbg}{RGB}{58, 26, 26}
\definecolor{freebg}{RGB}{26, 42, 58}
\definecolor{textlight}{RGB}{224, 224, 224}
\definecolor{textmuted}{RGB}{170, 170, 170}
\definecolor{borderopen}{RGB}{76, 175, 80}
\definecolor{borderclosed}{RGB}{231, 76, 60}
\definecolor{borderfree}{RGB}{52, 152, 219}

% Headers
\titleformat{\section}
  {\Large\bfseries\color{gold}}
  {}{0em}{}
\titleformat{\subsection}
  {\large\bfseries\color{gold}}
  {}{0em}{}

% Page style
\pagestyle{empty}

% Custom environments
\newenvironment{infobox}[2][darkbg]{
  \begin{framed}
  \setlength{\fboxsep}{12pt}
  \color{textlight}
}{
  \end{framed}
}

% Boxed section with colored left border
\newenvironment{policybox}[2]{
  \begin{framed}
  \setlength{\fboxsep}{12pt}
  \color{textlight}
  \noindent
  \textcolor{#2}{\rule{4pt}{\dimexpr\height+4pt\relax}}%
  \hspace{8pt}%
  \textbf{\textcolor{gold}{#1}}\par
  \vspace{4pt}
}{
  \end{framed}
}

\begin{document}

% ===== TITLE PAGE =====
\begin{titlepage}
  \centering
  \vspace*{2in}

  {\Huge\bfseries\color{gold} Fate's Edge}\\[0.2in]
  {\Large Third-Party Creator Policy}\\[0.1in]
  {\large Version 1.0 \textbullet\ June 2026}

  \vspace{1in}

  \rule{0.6\textwidth}{0.4pt}

  \vspace{0.5in}

  {\large A clear, fair guide to what you can do with Fate's Edge content}

  \vspace{0.5in}

  \rule{0.6\textwidth}{0.4pt}

  \vspace{1in}

  {\small \textcolor{textmuted}{
    \textit{``The coin that never spends is the one you don't remember taking.''}\\
    --- Serafine of the Velvet Touch
  }}

  \vfill

  {\small \textcolor{textmuted}{\textcopyright\ 2026 Nicholas A. Gasper}}
\end{titlepage}

% ===== CONTENT =====
\section*{Introduction}

Fate's Edge is a game built for community. The core mechanics—the dice system, Position/Effect, Boons and Story Beats, the TAGS magic system, and the character creation framework—are open for you to use, remix, and build upon. The world of the Amaranthine, its characters, its lore, and its adventures are the work of years of care. That stays close to the hearth.

This policy explains what you can do with Fate's Edge content, what you need permission for, and how to get it.

\section*{Part 1: What's Open (The SRD)}

The \textbf{Fate's Edge System Reference Document (SRD)} is licensed under \textbf{CC BY-NC-SA 4.0}.

\begin{policybox}{You May:}{borderopen}
  \begin{itemize}[leftmargin=*, itemsep=2pt]
    \item Use the mechanics in your own games, supplements, and adaptations
    \item Remix and adapt the core rules to fit your vision
    \item Share your work with the community—for free—under the same license
    \item Reference Fate's Edge mechanics in actual play streams, podcasts, and videos
    \item Create and sell content that \emph{uses} the SRD mechanics, as long as it does not incorporate extended material (see Part 2)
  \end{itemize}
\end{policybox}

\begin{policybox}{You Must:}{borderopen}
  \begin{itemize}[leftmargin=*, itemsep=2pt]
    \item Credit the creator (Nicholas A. Gasper) and link to the license
    \item Share your adaptations under the same CC BY-NC-SA 4.0 license
    \item Not use the SRD commercially in any product that competes with or incorporates extended Fate's Edge material
  \end{itemize}
\end{policybox}

\begin{policybox}{You May Not:}{borderopen}
  \begin{itemize}[leftmargin=*, itemsep=2pt]
    \item Claim the mechanics as your own—credit must be given
    \item Close the license—your adaptations must remain open
    \item Use Fate's Edge trademarks (logos, titles, character names, setting terms) without permission
  \end{itemize}
\end{policybox}

\section*{Part 2: What's Closed (Extended Content)}

The following are \textbf{All Rights Reserved} and require express written permission for any commercial use:

\begin{policybox}{Prohibited Without Permission:}{borderclosed}
  \begin{itemize}[leftmargin=*, itemsep=2pt]
    \item \textbf{Setting Lore:} The Amaranthine, its regions (Ecktoria, Acasia, Valewood, etc.), cultures, and history
    \item \textbf{Original Characters:} Named NPCs, Patrons, and iconic figures (e.g., Serafine of the Velvet Touch, the Gray Wanderer, Lady Isolde)
    \item \textbf{Adventures and Modules:} The Lantern at Dusk, Blood and Silk, Beyond Millhaven, The Brass Gate, and any published adventures
    \item \textbf{Proprietary Magic Systems:} Full patron entries, Invoker Rites, Cantor Songs, Runkeeper Codex content, and Summoner-specific mechanics as published
    \item \textbf{Artwork and Graphics:} All illustrations, maps, logos, and design elements
    \item \textbf{Faction and Organization Descriptions:} Detailed faction entries (e.g., The Vigil, the Black Banners, the Silk Vigil)
    \item \textbf{Region-Specific Generators:} Detailed generators and cultural mechanics as published in the Amaranthine Core
    \item \textbf{Original Prose and Framing Devices:} All narrative text, quotes, and in-character framing
  \end{itemize}
\end{policybox}

\section*{Part 3: Commercial Use Permission}

If you want to create and sell content that uses \textbf{extended material} (setting, characters, adventures, proprietary mechanics, etc.), you need permission.

\subsection*{How to Apply}

Send a request to \textbf{[PUBLISHER\_EMAIL]} with:

\begin{enumerate}[leftmargin=*, itemsep=2pt]
  \item Your name and contact information
  \item A description of your project (what it is, what it includes, who it's for)
  \item Which extended material you want to use (be specific)
  \item How you plan to distribute and price it
  \item A sample or prototype (if available)
\end{enumerate}

\subsection*{Terms}

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item Permission is granted on a \textbf{project-by-project basis}
  \item A \textbf{royalty} may be negotiated depending on the scope and use
  \item You must \textbf{clearly credit} Fate's Edge and Nicholas A. Gasper
  \item You may \textbf{not} imply official endorsement or affiliation
  \item \textbf{Final approval} rests with the author
\end{itemize}

\subsection*{Response Time}

Expect a response within \textbf{2–4 weeks}. If your project is time-sensitive, mention that in your request.

\subsection*{What We're Looking For}

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item Projects that \textbf{respect the tone and spirit} of Fate's Edge
  \item Work that is \textbf{well-written and professionally presented}
  \item Creators who are \textbf{open to collaboration and feedback}
  \item Content that \textbf{expands the world} rather than just copying it
\end{itemize}

\section*{Part 4: Trademark and Logo Usage}

Fate's Edge logos, titles, and branding are \textbf{trademarks} of Nicholas A. Gasper.

\subsection*{For Open Content (SRD-only)}

You may \textbf{not} use the Fate's Edge logo or title on your commercial or non-commercial work unless:

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item Your work is clearly marked as "A Fate's Edge Supplement" or "Fate's Edge Compatible"
  \item You include a disclaimer: "This product is not affiliated with or endorsed by the creator of Fate's Edge."
\end{itemize}

\subsection*{For Licensed Extended Content}

If you receive permission to use extended material, you \textbf{may} use the Fate's Edge logo, provided:

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item You follow the logo usage guidelines provided with your license
  \item You include a "Licensed Product" tagline or disclaimer as agreed
\end{itemize}

\section*{Part 5: Community Content (Free/Non-Commercial)}

You are \textbf{always welcome} to create and share free content for Fate's Edge.

\subsection*{You May (Without Permission):}

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item Write and share your own adventures, NPCs, or settings set in the Amaranthine
  \item Create fan art, fiction, or actual play content using Fate's Edge characters and locations
  \item Publish free supplements, rules hacks, and house rules using extended material
  \item Share your work on community forums, blogs, social media, or itch.io (free)
\end{itemize}

\subsection*{You Must:}

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item Clearly credit Fate's Edge and Nicholas A. Gasper
  \item Include a disclaimer: "This is unofficial content for Fate's Edge and is not affiliated with or endorsed by the creator."
  \item \textbf{Not sell} content that uses extended material without permission
\end{itemize}

\subsection*{We Love Seeing:}

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item New adventures set in the Amaranthine
  \item Character concepts and backgrounds
  \item House rules and system hacks
  \item Regional expansions and cultural explorations
  \item Actual play and fan fiction
\end{itemize}

Share your work! Tag us on social media and we'll help boost it.

\section*{Part 6: Reporting Violations}

If you see content that appears to violate this policy, please contact \textbf{[PUBLISHER\_EMAIL]}.

\section*{Summary Table}

\vspace{0.2in}

\begin{center}
  \begin{tabular}{>{\raggedright\arraybackslash}p{0.6\linewidth}>{\raggedright\arraybackslash}p{0.3\linewidth}}
    \toprule
    \textbf{What You Want to Do} & \textbf{Permission Needed?} \\
    \midrule
    Use SRD mechanics in your free supplement & No—SRD is open (CC BY-NC-SA) \\
    Use SRD mechanics in your paid supplement & No—SRD is open for commercial use \\
    Use the Amaranthine setting in a free adventure & No—free community content is welcome \\
    Use the Amaranthine setting in a paid adventure & Yes—commercial use requires permission \\
    Use named NPCs or Patrons in any commercial work & Yes—these are extended material \\
    Use Fate's Edge logo or title & Yes—trademark usage requires permission \\
    Create an actual play podcast & No—actual play is fine (credit appreciated!) \\
    Publish a free rules hack using SRD mechanics & No—SRD is open \\
    Publish a paid rules hack using extended material & Yes—commercial use requires permission \\
    \bottomrule
  \end{tabular}
\end{center}

\section*{Frequently Asked Questions}

\subsection*{Q: Can I publish my Fate's Edge adventure on DriveThruRPG?}

A: If it uses only SRD mechanics, yes. If it uses extended material (setting, characters, etc.), you need permission.

\subsection*{Q: Can I run a Fate's Edge campaign on Twitch or YouTube?}

A: Absolutely. No permission needed. We love seeing people play!

\subsection*{Q: Can I sell Fate's Edge merchandise (T-shirts, dice, etc.)?}

A: If it uses Fate's Edge logos, titles, or branding, you need permission. Generic "fantasy RPG" merchandise is fine.

\subsection*{Q: Can I translate the SRD into another language?}

A: Yes. The SRD is open for translation under the CC BY-NC-SA license. Translated versions must also be shared openly.

\subsection*{Q: Can I translate extended material (the setting, adventures, etc.)?}

A: Not without permission. Translations of extended material are derivative works and require explicit license.

\subsection*{Q: What if I want to publish a commercial supplement that uses the SRD but references the Amaranthine?}

A: If you reference the setting (e.g., "set in the city of Silkstrand"), that's extended material—you need permission. If you say "set in a fantasy city of canals and intrigue," that's fine—you're building your own world.

\section*{Version History}

\begin{itemize}[leftmargin=*, itemsep=2pt]
  \item \textbf{1.0} (June 2026): Initial release
\end{itemize}

\vspace{0.3in}

\noindent \textit{This policy is designed to be clear, fair, and encouraging. Fate's Edge is a game meant to be played, shared, and built upon. If you're unsure whether your project needs permission, just ask. We're happy to help.}

\section*{Contact}

For licensing inquiries or permission requests:

\vspace{0.1in}
\noindent \textbf{[PUBLISHER\_EMAIL]}

\vspace{0.3in}

\noindent\rule{\linewidth}{0.4pt}

\vspace{0.2in}
\noindent \textit{``The road is long. Walk it with people you trust—or at least people you owe. Both will carry you.''}\\
\hspace*{0.5in}--- The Gray Wanderer

\end{document}
