% Fate's Edge – TAGS Spellbook with Lookup Index
% Compile with XeLaTeX or LuaLaTeX

\documentclass[10pt,a5paper]{book}
\usepackage[top=1.5cm,bottom=1.5cm,left=1.2cm,right=1.2cm]{geometry}
\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}
\usepackage{titlesec}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}
\usepackage{multicol}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue}
\usepackage{array}
\usepackage{longtable}

\titleformat{\chapter}{\Huge\bfseries\scshape}{}{0em}{}
\titleformat{\section}{\Large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

\newtcolorbox{spellcard}[2][]{
    breakable,
    colback=gray!5,
    colframe=black!50,
    fonttitle=\bfseries\large,
    title=#2,
    enhanced,
    attach boxed title to top left={yshift=-3mm,xshift=3mm},
    boxed title style={colback=white},
    #1
}

\newcommand{\spell}[5]{ % name, tags, dv, effect, backlash
    \begin{spellcard}{#1}
    \textbf{TAGS:} #2 \\
    \textbf{DV:} #3 \\
    \textbf{Effect:} #4 \\
    \textbf{Backlash (1+ SB):} #5
    \end{spellcard}
    \vspace{0.5cm}
}

\newcommand{\dangerous}[1]{\textcolor{red}{\textbf{#1}}}
\newcommand{\tag}[1]{\textsc{#1}}

\begin{document}

\chapter*{TAGS Spellbook}
\tableofcontents

\section{Free Casting Rules}
A spell is built from \textbf{TAGS} (intent + target + tags).  
Base DV = \(1 + \text{number of TAGS}\).  
\dangerous{Dangerous TAGS} (marked \dangerous{⚠}) add \(+2\) instead of \(+1\). \\
\noindent\textbf{Casting roll:} \texttt{Attribute + Arcana} vs DV.  
Each \texttt{1} on the roll generates a Story Beat (SB) → Backlash (see table at end).

\newpage
\section{Tag Lookup \& Index}

\subsection*{How to read}
\begin{tabular}{@{} p{2.5cm} p{1cm} p{2.5cm} p{6cm} @{}}
\textbf{Tag} & \textbf{DV mod} & \textbf{Category} & \textbf{Effect summary} \\
\hline
\tag{Example} & +1 (+2 if dangerous) & Elemental & Brief mechanical effect \\
\end{tabular}
\vspace{0.5cm}

\begin{longtable}{@{} p{2.5cm} p{1.2cm} p{2.5cm} p{6cm} @{}}
\caption{TAG Reference}\\
\toprule
\textbf{Tag} & \textbf{DV mod} & \textbf{Category} & \textbf{Effect summary} \\
\midrule
\endfirsthead
\multicolumn{4}{c}{{\bfseries \tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Tag} & \textbf{DV mod} & \textbf{Category} & \textbf{Effect summary} \\
\midrule
\endhead
\bottomrule
\endfoot

% Elemental TAGS
\tag{Burning} & +1 & Elemental (Fire) & Ignite, heat, combustion, smoke \\
\tag{Freezing} & +1 & Elemental (Water) & Ice, slowing, brittle shatter, cold \\
\tag{Storm} & +1 & Elemental (Air) & Lightning, shock, arc, thunder \\
\tag{Stone} & +1 & Elemental (Earth) & Walls, spikes, tremors, armor \\
\tag{Wave} & +1 & Elemental (Water) & Crushing water, currents, pressure \\
\tag{Wind} & +1 & Elemental (Air) & Levitation, gusts, deflection, push/pull \\
\midrule

% Force TAGS
\tag{Force} & +1 & Force & Kinetic power, shields, blasts, telekinesis \\
\tag{Area} & +1 & Force & Cone, circle, corridor, zone effect \\
\tag{Strike} & +1 & Force & Single target precision \\
\tag{Wall} & +1 & Force & Barrier or blockade \\
\tag{Bind} & +1 & Force & Restrain, hold, suspend, entangle \\
\tag{Dispel} & +1 & Force & Suppress magic, unravel ongoing effects \\
\midrule

% Mind & Veil TAGS
\tag{Veil} & +1 & Mind/Illusion & Conceal, blur, illusion, silence \\
\tag{Scry} & +1 & Mind/Illusion & Reveal hidden, see distance, read traces \\
\tag{Memory} & +1 & Mind/Illusion & Erase, alter, restore memories \\
\tag{Command} & +1 & Mind/Illusion & Compel a short action (one word) \\
\tag{Fear} & +1 & Mind/Illusion & Panic, flee, break morale \\
\midrule

% Life & Body TAGS
\tag{Mend} & +1 & Life/Body & Close wounds, restore flesh, reduce Harm 1 \\
\tag{Purify} & +1 & Life/Body & Remove poison, corruption, disease \\
\tag{Strengthen} & +1 & Life/Body & Enhance body, armor, senses (temporary) \\
\tag{Waken} & +1 & Life/Body & Counter sleep, paralysis, stun \\
\tag{Beast} & +1 & Life/Body & Speak with or influence animals \\
\midrule

% Space & Motion (always dangerous)
\tag{Leap} & \dangerous{+2} & Space/Motion & Jump far, blink across short space (Near) \\
\tag{Fold} & \dangerous{+2} & Space/Motion & Short-range teleport, vanish–reappear (Far) \\
\tag{Gate} & \dangerous{+2} & Space/Motion & Long distance passage, open/close path \\
\tag{Gravity} & \dangerous{+2} & Space/Motion & Crush, lift, suspend, walk on walls/ceiling \\
\midrule

% Creation & Transformation (always dangerous)
\tag{Create} & \dangerous{+2} & Creation & Manifest mundane matter briefly (1 scene) \\
\tag{Summon} & \dangerous{+2} & Creation & Call a being or construct (see Summoning rules) \\
\tag{Transmute} & \dangerous{+2} & Creation & Turn one thing into another (temporary) \\
\tag{Animate} & \dangerous{+2} & Creation & Make objects act with intent (1 scene) \\
\midrule

% Additional Utility TAGS
\tag{Sense} & +1 & Utility & Detect presence of a named tag/element (Near) \\
\tag{Reveal} & +1 & Utility & Unveil hidden, glamoured, or invisible things \\
\tag{Light} & +1 & Utility & Create illumination (glow, torch‐bright) \\
\tag{Shadow} & +1 & Utility & Deepen darkness, hide edges, obscure \\
\tag{Silence} & +1 & Utility & Suppress sound in zone or on target \\
\tag{Protect} & +1 & Utility & Reduce/deflect next harm (Armor 1) \\
\midrule

% Reaction / Conditional (advanced)
\tag{Counter} & +1 & Reaction & Interrupt a casting/ritual in its window (requires held action) \\
\tag{Reflect} & \dangerous{+2} & Reaction & Turn next targeted effect back on its source \\
\tag{Store} & \dangerous{+2} & Utility & Bank 1–2 successes in a vessel (once) \\
\tag{Curse} & \dangerous{+2} & Affliction & Attach hostile tag/clock to target (requires Resolve save) \\
\tag{Bless} & +1 & Affliction & Grant favorable tag (luck, favor, ward-key) \\
\bottomrule
\end{longtable}

\section{Backlash Reference}
\begin{tcolorbox}[colback=red!5, colframe=red!50]
\begin{multicols}{2}
\textbf{Minor (1–2 SB)} \\
• Fatigue +1 \\
• –1 die on next roll \\
• Increase DV of next action by 1 \\
• Minor environmental effect (sparks, chill, scent) \\
• GM gains +1 SB for a thematic twist

\textbf{Major (3–4 SB)} \\
• Harm 1 (stress) \\
• Condition applied (Blinded, Silenced, Bound) \\
• Ward/barrier collapses \\
• Unwanted attention (Alarm clock advances) \\
• GM gains +2 SB to escalate the scene

\textbf{Catastrophic (5+ SB)} \\
• Harm 2 \\
• Permanent Scar or Mark \\
• Spell inverts (heal becomes harm, etc.) \\
• Patron or rival entity manifests \\
\end{multicols}
\end{tcolorbox}

\newpage
\section{Cantrips (DV 2–3)}
\spell{Ember Flick}{Burning, Strike}{2}{Ignite a small, unattended object within Close range; deal 1 Fatigue to a living target.}{Minor Backlash – spark jumps back; caster marks 1 Fatigue.}
\spell{Dampen Sound}{Veil, Silence}{2}{Muffle all sound from a single target or area (Near range) for one exchange.}{Minor Backlash – sound returns distorted; next Notice roll –1 die.}
\spell{Reed Walk}{Wind, Leap, Self}{3}{Leap up to Far distance in a straight line, ignoring opportunity attacks.}{Minor Backlash – land off‑balance; next action Desperate.}

\section{Combat Spells (DV 3–4)}
\spell{Storm Bolt}{Burning, Shock, Strike}{3}{A bolt of lightning strikes a single target at Near range. Deal Harm 1 (shock).}{Minor Backlash – residual arc jumps to nearest ally; they mark 1 Fatigue.}
\spell{Stone Shield}{Stone, Wall, Protect, Self}{4}{Raise a wall of stone (Close range, 1 zone) providing heavy cover for one scene; caster gains +1 Armor while adjacent.}{Major Backlash – wall crumbles prematurely; GM gains +2 SB to spend on environmental collapse.}
\spell{Grasping Roots}{Earth, Bind, Area}{3}{Roots erupt in a Near zone; all within must test Body+Athletics (DV 3) or become Entangled (lose next Move action).}{Minor Backlash – roots also snag an ally; they lose 1 step of Position.}

\section{Utility \& Exploration (DV 3–5)}
\spell{Memory Thread}{Memory, Veil, Touch}{3}{Touch a willing target to erase up to 10 minutes of their recent memory, or to implant a vivid false image.}{Major Backlash – caster loses a memory of equal value (GM chooses).}
\spell{Wind Guide}{Wind, Sense, Scry}{4}{Send a whisper into the wind; it will reach a specific person you know within 1 mile in one hour, carrying up to 10 words.}{Minor Backlash – the wind answers back with a cryptic omen (GM gains +1 SB).}
\spell{Shadow Fold}{Shadow, Veil, Area}{3}{Blanket a Near zone in supernatural darkness for one scene. Creatures without darkvision suffer –2 dice to attacks inside.}{Minor Backlash – darkness clings to the caster; they suffer –1 die to social rolls until washed in sunlight.}
\dangerous{\spell{Fate's Needle}{Fate, Strike, \dangerous{Divination}}{5}{Roll a pool of 3d10. Keep the highest die; that many successes automatically succeed on your next non‑magical action this scene.}{Major Backlash – the stolen fate demands payment; caster marks Harm 1 (stress) and begins a \textit{Fate Debt} clock [4].}}

\section{Ritual‑Grade Spells (DV 5+)}
\spell{Quick Gate}{\dangerous{Gate}, Leap, Self}{1+2+1+1 = 5}{Open a shimmering portal to a location you can see (Far range). You and one willing ally may step through as a Move action.}{Major Backlash – portal destabilises; you are shunted 1d3 zones in a random direction and take Harm 1 (dimensional shear).}
\spell{Animate Ward}{\dangerous{Animate}, Protect, Area, WARD}{1+2+1+1+1 = 6}{Give temporary sentience to a stationary ward or protective circle. For one scene, the ward can reposition itself (Move 1 zone per round) and automatically intercept one attack per round.}{Catastrophic Backlash – the ward becomes hostile, attacking the nearest creature for 1 round before burning out.}

\section{Building Your Own Spells}
\begin{enumerate}
    \item \textbf{Intent:} “I want to freeze the door shut.”
    \item \textbf{Choose TAGS:} \tag{Freeze}, \tag{Bind}, \tag{Wall}, \tag{Touch}.
    \item \textbf{Count TAGS:} 4 → Base DV = 1+4 = 5.
    \item \textbf{Adjust for scope:} Large door DV remains 5. Entire fortress gate? DV +2.
    \item \textbf{Roll:} Wits+Arcana vs DV 5. Each 1 = SB.
    \item \textbf{Apply Backlash per table.}
\end{enumerate}
\vspace{0.5cm}
\textit{“Channel bold, weave true, and may the Story Beats be ever in your favour.”}

\end{document}
