\documentclass[10pt]{article}
\usepackage[letterpaper,margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{graphicx}

% Color definitions
\definecolor{charblue}{RGB}{30,80,120}
\definecolor{lightgray}{RGB}{240,240,240}
\definecolor{darkgray}{RGB}{60,60,60}

% Box styles
\tcbuselibrary{skins,breakable}
\newtcolorbox{infobox}[1][]{colback=lightgray,colframe=charblue,title=#1,fonttitle=\bfseries,sharp corners,breakable}
\newtcolorbox{warningbox}{colback=red!5,colframe=red!70!black,sharp corners,breakable}
\newtcolorbox{tipbox}{colback=green!5,colframe=green!70!black,sharp corners,breakable}

\begin{document}

% --- Title Page ---
\begin{center}
\vspace*{2cm}
{\Huge \textbf{Fate's Edge}}

\vspace{0.5cm}
{\huge \textbf{Read Me First}}

\vspace{0.3cm}
{\Large A Guide to Navigating the Deluxe Package}

\vspace{1cm}
{\large Nicholas A. Gasper}

\vspace{0.3cm}
{\large July 1, 2026}

\vspace{2cm}
{\itshape
``The road does not promise a safe journey---only a true one.'' \\
--- The Gray Wanderer
}

\vspace{2cm}
\rule{0.4\textwidth}{0.5pt}

\vspace{1cm}
{\large \textbf{Start Here. Then Read What You Need.}}
\end{center}

\newpage

% --- SECTION 1: WELCOME ---
\section*{Welcome to Fate's Edge}

You have just opened a \textbf{deluxe package} of game documents. This is not a single book---it is a \textbf{library}. The material covers:

\begin{itemize}
\item Core rules (available in two formats: a \textit{teaching} guide and a \textit{reference} document)
\item A complete setting (the Amaranthine Sea and its regions)
\item Player-facing material (character creation, magic, advancement)
\item Game Master material (campaign design, timers, factions)
\item Adventures (a complete campaign arc)
\item Quick-reference tools (GM Screen, Player Cyclopedia)
\end{itemize}

\noindent
This is a lot of material. You do \textbf{not} need to read all of it.

This document tells you where to start based on \textbf{who you are} and \textbf{what you need}.

\vspace{0.5cm}
\begin{tipbox}
\textbf{The Golden Rule:} You own the game. The game does not own you. Read what you need, skip what you don't. You can always come back for more.
\end{tipbox}

\newpage

% --- SECTION 2: DOCUMENT MAP ---
\section*{The Document Map}

The Fate's Edge deluxe package consists of the following documents:

\vspace{0.3cm}
\begin{center}
\begin{tabular}{|p{3.5cm}|p{8cm}|}
\hline
\textbf{Document} & \textbf{What It's For} \\
\hline
\textit{Read Me First} (this document) & Your entry point. Tells you where to go next. \\
\hline
\textbf{Player Cyclopedia} & One-page quick reference for players. Keep at the table. \\
\hline
\textbf{Essential GM Screen} & One-page quick reference for GMs. Keep behind the screen. \\
\hline
\textbf{Essentials Guide} & The core rules, taught through examples. Start here if you're new. \\
\hline
\textbf{SRD (System Reference Document)} & The core rules, organized as a reference. Use when you need to look something up. \\
\hline
\textbf{Player's Guide} & Character creation, magic, advancement, and player-facing material. \\
\hline
\textbf{GM Guide} & Campaign design, timers, factions, assets, followers, and advanced techniques. \\
\hline
\textbf{Amaranthine Sea (Setting)} & The world. Regions, cultures, cartomancy, and the Ninth Taboo. \\
\hline
\textbf{Blood and Silk (Prequel)} & The fall of the Silk Coin. A tragic one-shot. \\
\hline
\textbf{Blood and Silk I: The Lantern at Dusk} & Starter adventure. Village defense. \\
\hline
\textbf{Blood and Silk II: Beyond Millhaven} & Sequel. Negotiation and mediation. \\
\hline
\textbf{Blood and Silk III: The Brass Gate} & Finale. Heist and curse-breaking. \\
\hline
\textbf{Character Sheet Booklet} & Pre-generated characters and build guidance. \\
\hline
\end{tabular}
\end{center}

\newpage

% --- SECTION 3: YOUR ROLE ---
\section*{What to Read Based on Your Role}

\subsection*{If You Are a Player}

\begin{enumerate}
\item \textbf{Read First:} \textit{Player Cyclopedia} (1 page). This gives you the core rules in under 60 seconds.
\item \textbf{Read Next:} \textit{Player's Guide}, Chapters 1-3 (Quick Start, Character Creation Summary, Core Mechanics). This gives you everything you need to build and run a character.
\item \textbf{If You Want Magic:} \textit{Player's Guide}, Chapter 4 (Magic and Special Abilities) or Chapter 9 (Your First Magical Character).
\item \textbf{If You Want Setting:} \textit{Amaranthine Sea}, Chapter 2 (On Obligation and Trade) and the region your campaign is set in.
\item \textbf{You Do Not Need:} The SRD (it's reference, not teaching), the GM Guide (that's for your GM), or the adventures (unless you want to peek).
\end{enumerate}

\begin{tipbox}
\textbf{Player Tip:} You can play a complete, effective character with just Attributes and Skills. Talents are optional flavor. Start simple, add complexity as you learn.
\end{tipbox}

\subsection*{If You Are a Game Master}

\begin{enumerate}
\item \textbf{Read First:} \textit{Essential GM Screen} (1 page). This gives you the core resolution loop at a glance.
\item \textbf{Read Next:} \textit{Essentials Guide}. This teaches the core rules through examples. You need to know how the system works before you run it.
\item \textbf{Read Next:} \textit{GM Guide}, Chapters 1-3 (Foundations, Core Procedures, Managing Resources). This gives you the practical tools for running the game.
\item \textbf{If You Want to Run an Adventure:} Start with \textit{Blood and Silk I: The Lantern at Dusk}. It's a complete starter adventure with pre-generated characters.
\item \textbf{If You Want Setting:} \textit{Amaranthine Sea}, the region your campaign is set in. Acasia is recommended for new GMs.
\item \textbf{If You Want Advanced Techniques:} \textit{GM Guide}, Chapters 10-12 (Campaigns, Timers, Consequences; Tier IV-V; Advanced Techniques).
\item \textbf{You Do Not Need:} The SRD (you have the Essentials Guide), the Player's Guide (unless you want to confirm something), or the adventures beyond the one you're running.
\end{enumerate}

\begin{warningbox}
\textbf{GM Warning:} Do not try to read everything before your first session. You only need the Essentials Guide, the GM Guide Chapters 1-3, and one adventure. The rest is optional depth.
\end{warningbox}

\newpage

% --- SECTION 4: CORE RULES SUMMARY ---
\section*{The Core Rules in 2 Pages}

This is the essential resolution system. Every other rule builds on this foundation.

\subsection*{The 90-Second Play Loop}

\begin{enumerate}
\item \textbf{Declare intent and approach.} What do you want to achieve, and how? (e.g., ``I sneak past the guard using Wits + Stealth.'')
\item \textbf{GM sets DV and Position.} Difficulty Value (2-5+) and Position (Dominant/Controlled/Desperate).
\item \textbf{Roll dice.} Roll Attribute + Skill d10s.
\item \textbf{Count successes.} Each die showing 6+ = 1 success. Each 10 = 2 successes.
\item \textbf{Count Story Beats.} Each die showing 1 = 1 SB for the GM.
\item \textbf{Apply the Outcome Matrix.}
\item \textbf{GM spends Story Beats.} Complications are introduced.
\item \textbf{Player marks resources.} Boons, Fatigue, Harm, etc.
\end{enumerate}

\subsection*{Difficulty Value (DV)}

\begin{center}
\begin{tabular}{|c|l|}
\hline
\textbf{DV} & \textbf{When to Use} \\
\hline
2 & Routine, almost guaranteed \\
3 & Pressured, mild opposition (default) \\
4 & Hard, serious resistance \\
5+ & Extreme, dramatic gamble \\
\hline
\end{tabular}
\end{center}

\subsection*{Position (Risk)}

\begin{center}
\begin{tabular}{|c|l|c|}
\hline
\textbf{Position} & \textbf{Narrative Frame} & \textbf{Mechanical Effect} \\
\hline
Dominant & You press your advantage & Re-roll one failure \\
Controlled & Balanced, normal risk & No re-rolls \\
Desperate & You act under duress & Re-roll one success \\
\hline
\end{tabular}
\end{center}

You may spend 1 Boon to improve Position by one step before rolling.

\subsection*{The Outcome Matrix}

\begin{center}
\begin{tabular}{|c|l|}
\hline
\textbf{Roll Result} & \textbf{What Happens} \\
\hline
Successes $\geq$ DV, SB = 0 & Clean Success: intent achieved, no complication \\
Successes $\geq$ DV, SB $>$ 0 & Success with SB: intent achieved; GM spends SB \\
0 $<$ Successes $<$ DV & Partial: progress made; player gains 1 Boon \\
Successes = 0 & Miss: no progress; player gains 2 Boons; GM escalates \\
\hline
\end{tabular}
\end{center}

\newpage

\subsection*{Resources at a Glance}

\begin{center}
\begin{tabular}{|p{4cm}|p{9cm}|}
\hline
\textbf{Resource} & \textbf{Summary} \\
\hline
\textbf{Boons} & Player currency. Earn on Partial (1) or Miss (2). Max 5 held. Spend to re-roll a die, improve Position, activate an Asset, or convert 2 $\to$ 1 XP (once/session). \\
\hline
\textbf{Story Beats (SB)} & GM currency. Earn every time a player rolls a 1. Spend to introduce complications: 1 SB = minor, 2 SB = moderate, 3+ SB = major. \\
\hline
\textbf{Fatigue} & Track size = Body. Each point worsens Position. When track fills $\to$ Harm +1, clear all Fatigue. Short rest clears 2; full night clears all. \\
\hline
\textbf{Harm} & Level 1: $-1$ die on related actions. Level 2: $-1$ die on most actions. Level 3: incapacitated or dying. \\
\hline
\textbf{Armor Conversion} & Armor converts Harm to Fatigue. Minimum 1 Fatigue per hit. Taking Harm clears all existing Fatigue (the ``roller-coaster'' effect). \\
\hline
\end{tabular}
\end{center}

\subsection*{Combat Quick Reference}

\begin{itemize}
\item \textbf{Turn:} 1 Action + 1 Move
\item \textbf{Move:} Shift one range band (Close $\leftrightarrow$ Near $\leftrightarrow$ Far)
\item \textbf{Opportunity Attack:} Leaving Close range without Disengage provokes one free melee attack.
\item \textbf{Sidestep Strike:} Move one band as part of a melee attack; worsen Position by one step.
\item \textbf{Shove:} Opposed Body + Athletics; push target one band.
\item \textbf{Grapple:} Opposed Body + Melee; both Engaged; target suffers $-1$ die.
\end{itemize}

\subsection*{Golden Rule}

\begin{tipbox}
\textbf{When in doubt, make the choice that serves the story.} If you cannot remember a rule, default to DV 3, Controlled Position, and let the dice fall.
\end{tipbox}

\newpage

% --- SECTION 5: SAMPLE CHARACTER ---
\section*{Sample Character}

Here is a complete starting character (32 XP) to show how the rules come together.

\subsection*{Lyra the Scholar}

\begin{infobox}[Character Sheet]
\textbf{Concept:} A former Academy scholar, expelled for forbidden research.

\textbf{Attributes:} \\
Wits 3 (15 XP), Spirit 2 (6 XP) \\
Body 1, Presence 1

\textbf{Skills:} \\
Lore 2 (6 XP), Investigation 1 (2 XP), Insight 1 (2 XP)

\textbf{Talents:} \\
Lorekeeper (2 XP) --- Once per session, recall obscure history without rolling.

\textbf{Asset:} \\
Scholar's Satchel --- Contains books, reagents, a small telescope.

\textbf{Bonds:} \\
``We both lost our place in Silkstrand.'' (with another PC)

\textbf{Total XP:} 31 (bank 1)

\textbf{Typical Dice Pools:} \\
Wits 3 + Lore 2 = 5 dice to recall ancient history. \\
Wits 3 + Investigation 1 = 4 dice to find clues. \\
Spirit 2 + Lore 2 = 4 dice to resist fear.
\end{infobox}

\subsection*{How to Play Lyra}

\begin{itemize}
\item Lyra is a \textbf{knowledge specialist}. She rolls 5 dice for Lore checks, making her excellent at research, identifying relics, and understanding local customs.
\item She has \textbf{no combat skills}. In a fight, she should use her Lorekeeper talent to recall a weakness or use her Scholar's Satchel to provide tools for allies.
\item Her bond with another PC gives her a \textbf{once-per-session Boon} when she acts on it. Example: ``I remember that time we escaped the Academy together. I won't abandon you now.''
\item She is a \textbf{complete character} with zero magical abilities. Her power is in knowledge, not spells.
\end{itemize}

\newpage

% --- SECTION 6: SAMPLE SCENE ---
\section*{Sample Scene of Play}

To show the core loop in action:

\subsection*{Setup}

Lyra and her party are in a ruined temple. They need to cross a collapsed floor to reach an ancient artifact. The floor is unstable.

\subsection*{Play}

\textbf{GM:} ``The floor groans beneath you. You can see the artifact on a pedestal 20 feet away, but the stonework looks like it could collapse at any moment. What do you do?''

\textbf{Lyra's Player:} ``I want to find a safe path across. I'll use my knowledge of old architecture to identify which stones are load-bearing. That's Wits + Lore, right?''

\textbf{GM:} ``Yes. This is a \textbf{Controlled} Position---you have time to study the floor. The Difficulty Value is \textbf{3}---it's a tricky old ruin, but you're a scholar.''

\textbf{Lyra's Player:} ``Rolling Wits 3 + Lore 2 = 5 dice.''

\noindent
(Rolls: 7, 6, 3, 2, 1)

\textbf{GM:} ``Two successes (7 and 6). That's a \textbf{Partial} success. You find the safe path, but you also dislodge a stone. You gain 1 Boon. And that 1 gives me a Story Beat.''

\textbf{Lyra's Player:} ``I gain a Boon. What happens?''

\textbf{GM:} ``I spend the Story Beat. The stone you dislodged falls into the chamber below---and something stirs. You hear a low growl from the darkness. You have the safe path, but you've also attracted attention. What do you do?''

\subsection*{What Just Happened}

\begin{itemize}
\item Lyra's player declared intent and approach (Wits + Lore).
\item GM set Position (Controlled) and DV (3).
\item Lyra rolled 5 dice and got 2 successes.
\item The result was a Partial (successes $<$ DV)---progress was made, but at a cost.
\item Lyra gained 1 Boon for the Partial.
\item The 1 on the die gave the GM 1 Story Beat.
\item The GM spent the SB to introduce a complication: something stirred in the darkness.
\item The scene continues with new pressure.
\end{itemize}

\vspace{0.5cm}
\begin{tipbox}
\textbf{Key Lesson:} A Partial is not failure. It moves the story forward. Even on a success, the game can introduce complications. Every roll matters.
\end{tipbox}

\newpage

% --- SECTION 7: WHAT TO READ NEXT ---
\section*{What to Read Next}

Based on your role, here is your recommended reading order:

\subsection*{For Players}

\begin{enumerate}
\item \textit{Player Cyclopedia} --- 1 page, 2 minutes.
\item \textit{Player's Guide}, Chapter 1 (Quick Start) --- 5 minutes.
\item \textit{Player's Guide}, Chapter 2 (Character Creation Summary) --- 15 minutes.
\item \textit{Player's Guide}, Chapter 3 (Core Mechanics) --- 20 minutes.
\item \textit{Player's Guide}, Chapter 4 (Magic, if applicable) --- 15 minutes.
\item \textit{Player's Guide}, Chapter 9 (Your First Magical Character, if applicable) --- 10 minutes.
\end{enumerate}

\textbf{Total:} $\sim$1 hour, and you're ready to play.

\subsection*{For Game Masters}

\begin{enumerate}
\item \textit{Essential GM Screen} --- 1 page, 2 minutes.
\item \textit{Essentials Guide} --- 30 minutes. (You need to know the rules to run them.)
\item \textit{GM Guide}, Chapters 1-3 (Foundations, Core Procedures, Managing Resources) --- 45 minutes.
\item \textit{Blood and Silk I: The Lantern at Dusk} (the starter adventure) --- 20 minutes.
\item \textit{Amaranthine Sea}, the region your campaign is set in (Acasia recommended for new GMs) --- 15 minutes.
\end{enumerate}

\textbf{Total:} $\sim$2 hours, and you're ready to run your first session.

\subsection*{For Veterans (Experience with Other RPGs)}

\begin{enumerate}
\item \textit{Player Cyclopedia} or \textit{Essential GM Screen} (1 page).
\item \textit{Essentials Guide}, Chapter 2 (Core Mechanics) --- 15 minutes.
\item \textit{Essentials Guide}, Chapter 3 (Character Creation) --- 15 minutes.
\item \textit{GM Guide}, Chapter 1 (Introduction: The Weight of Choice) --- 10 minutes.
\end{enumerate}

\textbf{Total:} $\sim$40 minutes, and you'll understand how Fate's Edge differs from what you know.

\subsection*{For Magic Users}

\begin{enumerate}
\item \textit{Player's Guide}, Chapter 9 (Your First Magical Character) --- 15 minutes.
\item \textit{Player's Guide}, Chapter 4 (Magic and Special Abilities) --- 30 minutes.
\item \textit{Player's Guide}, Chapter 13 (World Powers, Patron section) --- 20 minutes.
\end{enumerate}

\vspace{0.5cm}
\begin{infobox}[Recommended First Patron]
For your first magical character, choose \textbf{The Traveler}. Its Rites are forgiving, Obligation costs are low, and it's thematically flexible for any character concept.
\end{infobox}

\newpage

% --- SECTION 8: QUICK TIPS ---
\section*{Quick Tips for New Players}

\begin{itemize}
\item \textbf{You don't need Talents to be effective.} Attributes and Skills do the heavy lifting. Talents are flavor, not viability.
\item \textbf{Spend your Boons.} Hoarding them for XP conversion starves the table of drama. A re-rolled failure or an improved Position creates a memorable moment.
\item \textbf{Embrace failure.} A Miss is never ``nothing happens.'' It's an opportunity for the GM to escalate, for your character to learn, and for the party to adapt. You gain 2 Boons when you Miss.
\item \textbf{Flavor is free.} Describe your actions with detail. The GM may reward you with Position, Effect, or a Boon.
\item \textbf{You are a co-creator.} The GM frames the pressure; you drive the fiction. You are not trying to ``beat'' the game.
\item \textbf{Trust the system.} The core loop is robust enough to carry any scene. If you forget a rule, fall back on the Golden Rule: \textit{make the choice that serves the story.}
\end{itemize}

\subsection*{Quick Tips for New GMs}

\begin{itemize}
\item \textbf{Default to DV 3, Position Controlled.} You can look up the rule after the session.
\item \textbf{Spend Story Beats early and often.} Unused pressure is wasted momentum.
\item \textbf{Fail forward.} On a Miss, never say ``nothing happens.'' Always introduce a new complication or advance a timer.
\item \textbf{Use the pre-generated characters.} The Character Sheet Booklet has complete characters for your first session.
\item \textbf{The Three-Timer Rule:} Keep at most three active timers in a scene. More creates cognitive overload.
\item \textbf{You are playing the game too.} Your job is not to defeat the players. Your job is to make the story memorable.
\end{itemize}

\newpage

% --- SECTION 9: FINAL ADVICE ---
\section*{Final Advice}

\begin{center}
\itshape
``The road does not promise a safe journey---only a true one.''

\vspace{0.5cm}
--- The Gray Wanderer
\end{center}

\vspace{1cm}

You have everything you need to play Fate's Edge. You do not need to read every document. You do not need to memorize every rule. You need:

\begin{enumerate}
\item A clear understanding of the core resolution loop.
\item A character (or a group of characters).
\item A GM who knows how to set Position, DV, and spend Story Beats.
\item A willingness to embrace failure as fuel for drama.
\end{enumerate}

\textbf{Start simple.} Use the pre-generated characters. Run \textit{The Lantern at Dusk}. Add complexity as you grow comfortable. The road is long, and it does not demand that you run before you walk.

\vspace{0.5cm}

Now close this document. Pick up the one that matches your role. And let the dice fall.

\vspace{1cm}

\begin{center}
\rule{0.6\textwidth}{0.5pt}

\textit{``The coin that never spends is the one you don't remember taking.''} \\
--- Serafine of the Velvet Touch

\rule{0.6\textwidth}{0.5pt}
\end{center}

\input{copyright}
\end{document}