\documentclass[11pt]{book}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{tcolorbox}

\geometry{margin=1in}
\setlength{\parskip}{6pt}

\begin{document}

\title{Fate's Edge: Player's Guide Primer}
\author{}
\date{}
\maketitle

\tableofcontents

\chapter{Welcome to Fate's Edge}
Fate's Edge is a world where choices ripple outward, magic is perilous, and cultures define more than appearance—they shape destiny. 
This primer offers the foundation: lands, tongues, gods, and peoples. It also introduces the three XP paths that shape characters in play.

\chapter{The Northern Shore \& Amaranthine Sea}
The heart of adventure lies along the Amaranthine Sea and its northern shore, a mosaic of kingdoms, republics, and wilderness.

\section{Major Regions}
\subsection*{Post–Utar / Upper Amaranthine Basin — “Shards of an Empire”}
\textbf{Tagline:} River thrones, hill banners, and city-states stitched together by roads older than their rulers.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Successor realms ring the Amaranthine and Belworth systems: old Utar law persists in ledgers, not loyalties.
  \item Two great confessions—Everflame and Light—compete with local cults and dwarven compacts.
  \item Roads, ferries, and coin-houses matter as much as castles; mercenary companies thrive.
\end{itemize}

\begin{description}[leftmargin=2.2cm]
  \item[Ecktoria] Rump of the Utar heartland: gladiator arenas, Everflame processions, marble in need of mortar. A wealthy capital city rides habits of empire while petty principalities fill the hinterland. Exports ceremony and coin, imports grain and peace.

  \item[Acasia] Failed province of ruined prefectures and opportunist crowns; strongmen rule mile by mile. \emph{Silkstrand} is the bright anomaly—oligarch merchants, bravos, and blue-water trade. Caravans hug safe corridors; everyone else pays someone.

  \item[Vhasia] Old union of blood-houses and dukedoms; pride hardens into rivalry. Lence keeps the royal fiction alive; Vhaston, Eastwell, and Ballafield pull in different directions. Oath, marriage, and musters remain the coin of governance.

  \item[Viterra] Highlands, fens, and dales under a lawful temperament and long memory. River tolls are counted to the nail; the King’s/Queen’s Highway customs are sacred. Knights of the Dawn keep roads honest; counting houses keep everyone else so.

  \item[Ubral] Foothill clans and hill dwarves knit by feud law and fast marches. Hit-and-run war is a craft; hospitality is binding once given. Trade runs in cattle, iron blooms, and guides who know every sheep-track and scree path.

  \item[Thepyrgos] Astroegro’s last lantern: universities, orators, ateliers, and archives amid kraken-haunted lore. High-elf enclaves maintain salons and schools; harbors hum with debate as much as ships. Beauty, rhetoric, and contracts share a bench.

  \item[Mistlands] Dwarven “protectorate” of peat farms and fog, where waystones and ward-mounds hold back stranger things. Payden’s Port is the hinge; fae incursions are a seasonal line item. Human steads prosper—inside agreed boundaries.

  \item[Theona] Three tight islands of hedgerows, marble cuts, and yeoman plots; duchies argue quietly while fields work loudly. Claims from the mainland wax and wane; parish charters outlast dukes. Grain, stone, and stubbornness export well.

  \item[Kahfagia] Maritime oligarchy on the western quarter: army, navy, marines, clergy, civics, and privateers share the helm. Tempest rites bless pilots and warlocks alike; charter ports project power. Finished goods out, prize law in.

  \item[Galanina \& Fharan Lands] Cedar coasts and caravan capitals: Sidhi courts, Fharan trade cities, and shrine networks that make deserts crossable. Silk and spice routes braid cultures; litigies and ledgers travel with caravans.

  \item[Haayr Peninsula] Mountain-ridged hinge between seas; palimpsest of empires. Ancient Hyro stones, Ashaani ruins, and current Kuvani prefectures overlap. Whoever holds the passes writes the tariffs—until the next holder does.
\end{description}

\section{Travelers' Map (Optional Placeholder)}

\noindent
\textit{This section is reserved for a visual map of the Amaranthine Sea region and surrounding lands. Until the map is finalized and included with:}  
\begin{center}
  \texttt{\textbackslash includegraphics[width=\textbackslash textwidth]\{map\_fatesedge.pdf\}}
\end{center}

\subsection*{Orientation Notes}
Until the map is prepared, use the following compass-points summary to orient play:  

\begin{description}[leftmargin=2cm]
  \item[North] Mistlands and Payden’s Port; dwarf protectorates and Fae-haunted fogs.  
  \item[East] Valewood, Amedell, and the Alberriden Sea; homeland of Wood Elves and Halflings.  
  \item[South] Amaranthine Sea and the broken provinces of Utar; Ecktoria, Vhasia, and Viterra dominate.  
  \item[West] Kahfagia across the Titan’s Throat, naval empire of storms and krakens.  
\end{description}

\subsection*{Major Routes}
\begin{itemize}
  \item \textbf{The Belworth Corridor} — Main artery between Vhasia and Viterra, heavily fortified.  
  \item \textbf{The Violet Steppes} — Wide-open plains north of Utar, cradle of Ykrul incursions.  
  \item \textbf{The Way of Silk} — Southern caravan trade linking Fhara, Kuvani, and Dhahara to northern ports.  
  \item \textbf{The Titan’s Throat} — The narrow strait where the Amaranthine meets the Great Sea; lifeline for Kahfagian fleets.  
\end{itemize}

\subsection*{Use at the Table}
\begin{itemize}
  \item \textbf{As a GM:} Drop pins for encounters, migrations, or rumors to create a sense of travel.  
  \item \textbf{As a Player:} Use the cardinal notes above to keep track of where your character has journeyed.  
  \item \textbf{At Character Creation:} Mark one home region and one place of past travel to immediately ground your PC in the world.  
\end{itemize}


\chapter{Languages of the Lands}
Language is both bridge and barrier. Most folk speak \textbf{Common}, a dwarven-forged trade tongue. 
Yet each culture prizes its own tongue:
\begin{itemize}
  \item \textbf{Low Utaran} — the speech of peasants and soldiers.  
  \item \textbf{High Utaran} — the scholar’s and priest’s tongue.  
  \item \textbf{Vhasian Dialects} — Lence’s courtly form vs. Vhaston’s trader’s patter.  
  \item \textbf{Viterran Plain Speech} — blunt, pragmatic, precise.  
  \item \textbf{Dwarven (Aeler)} — guttural, clipped, rich in trade terms.  
  \item \textbf{Elven (Lethai)} — ancient, fluid, contextual.  
  \item \textbf{Tulkani Tongue} — lilting, mobile, woven with shadow-cant.  
  \item \textbf{Kuvani Speech} — sharp, consonant-heavy, linked to steppe songs.  
\end{itemize}

\section{Far-Off Lands (Brief Gazetteer)}

\\subsection*{Akilan — “Continent of Sun, Stone, and River”}
\textbf{Tagline:} Cradle of empires, where river silt feeds scholar-kings and savanna caravans carry the weight of gold.\\
\textbf{Snapshot:}
\begin{itemize}
  \item A vast continent of monsoon coasts, river-fed empires, and savanna caravan routes linking the inland riches to blue-water ports.
  \item Power centers range from the bureaucratic machine of Oshiira and the divine mandate of Taharka to the mercantile hinges of Sekogo and the stone-kept secrets of the Ngomebe.
  \item Exports: Gold, spices, hardwoods, ivory, trained scribes, and monsoon pilots. Imports: Wrought steel, northern grain, books, and mercenary contracts.
  \item Languages: Oshiiran, Taharkan, Sekogo trade-pidgin, Ngombe stone-cant, and ancient Ashaani liturgical scripts.
  \item Hooks: An Oshiiran prefect hires guards for a census in the restless Crimson Basin; a Taharkan convoy master needs discreet muscle to bypass a Sekogo quay syndicate's ``weighing fee''; a Ngombe mason-guild offers strong coin for the recovery of a stolen resonance-crystal from a sunken temple.
\end{itemize}

\begin{center}
\begin{tabular}{|p{3cm}|p{3.5cm}|p{3.5cm}|p{5.5cm}|}
\hline
\textbf{Nation} & \textbf{Government/Identity} & \textbf{Exports} & \textbf{Adventure Hook} \\
\hline
\textbf{Oshiira} & Scholar-bureaucracy; riverine prefectures ruled by archivists and census-takers. & Scribes, law codes, jade seals. & A missing tax-roll imperils a governor’s mandate. \\
\hline
\textbf{Taharka} & Theocratic dynasty claiming divine sun-mandate; convoy masters and priest-kings. & Gold dust, ritual bronze, caravan oxen. & A rival sun-priest raises a heretical banner; adventurers hired to suppress it. \\
\hline
\textbf{Sekogo} & Mercantile hinge; coastal syndicates and guild councils dominate trade. & Salt, kola, indigo, mercantile services. & A syndicate clerk offers coin if smugglers vanish before tomorrow’s ledger is read. \\
\hline
\textbf{Ngomebe} & Mountain stone-keepers; resonance-stone masons and secret archivists. & Resonance crystals, cut stone, elephant ivory. & A fallen vault floods a valley with strange harmonics; escorts needed to reseal it. \\
\hline
\end{tabular}
\end{center}

\subsection*{Sihai — “Thousand Terraces, Ten Thousand Seals”}
\textbf{Tagline:} River dynasties, scholar-magistrates, and calendars that bind empires.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Heartland river plains ringed by mountains; irrigation guilds and terrace clans feed sprawling capitals.
  \item Bureaucracy of examination-seals; mandarins arbitrate clan feuds, salt monopolies, and spirit-rights.
  \item Exports: tea, silk, paper, porcelain, ink-stones. Imports: horses, fine ore, foreign mercenaries.
  \item Languages: High Sihai (court/ritual), Low Sihai (market); Common on the frontier.
  \item Hooks: A counterfeit seal-ring undermines a prefecture; a dragon-flood threatens the grain fleet unless taboo stones are moved.
\end{itemize}

\subsection*{Dhahara — “Peninsula of Sky and Memory”}
\textbf{Tagline:} Monsoon coasts and dust kingdoms where the god of the open sky still judges men by their banners.\\
\textbf{Snapshot:}
\begin{itemize}
  \item A long peninsula of river belts, thorn-steppe, and desert basins; caravan roads link upland passes to warm-water ports.
  \item \textbf{North}: the \emph{Himdal Marches}—Aeler-administered marches where old Dhaharan clans and ancient Kuvani blood intermix under fortress law. 
  \item \textbf{South}: Kahfagian charter-ports and “concessions” dominate deep-harbor trade; local courts retain inland rule but answer to quay ledgers.
  \item The majority faith venerates \textbf{Dhahar}, Lord of the Sky and War; coastal cities layer Tempest rites for safe harbors; caravan shrines keep the Way of Silk/Spice.
  \item The \textbf{Temple of Sealing} (old Dhaharan trust) and its sundered orders: a public fraternal lineage that treats the relic as legend, and a hardline cloister that keeps the true warnings.
  \item Exports: war-mounts, fine cotton, dyes, pepper, frankincense/myrrh resins, carved jade, caravan services. Imports: northern steel, coin-credit, ship timbers, learned clerks.
  \item Languages: Dhaharan (court and caravan), Himdali Aeler (march law), Kahfagian port-cant; Common on the roads.
  \item Hooks: A sky-feast tourney turns bloody after a banner is defamed; a concession-port freezes a convoy over “storm-damage fees”; a Himdal prefect seeks neutral arbiters for a clan water-rights moot.
\end{itemize}

\begin{center}
\begin{tabular}{|p{3.4cm}|p{4.1cm}|p{4.2cm}|p{5.2cm}|}
\hline
\textbf{Polity / Region} & \textbf{Government / Identity} & \textbf{Known For} & \textbf{Adventure Hook} \\
\hline
\textbf{Himdal Marches} (Northern uplands) & Aeler praefects and march-lords over mixed Dhaharan--Kuvani clans; fortress law, road tolls, census writs. & Border horse, hill-forts, pass control, levy-archers. & A census stone goes missing; three clans blame each other—recover it before the prefect levies a punitive tithe. \\
\hline
\textbf{Upper Dhahara} (Interior basins) & Clan emirs and sultanates bound by sky-oaths; water-right councils arbitrate grazing and wells. & Pepper gardens, incense groves, banner-tourneys, caravan guards. & A sky-oath duel is fixed; prove it without shaming both houses and starting a banner-feud. \\
\hline
\textbf{Lower Dhahara} (Monsoon belt) & City councils under hereditary malikates; Dhahar rites blended with harbor offerings. & Cotton looms, dye-vats, river-quays, monsoon pilots. & Monsoon charts are forged; expose the quay-syndicate before a fleet founders on a lee shore. \\
\hline
\textbf{Kahfagian Concessions} (Deep ports) & Charter harbormasters answer to Kahfagia’s oligarchy; local courts kept at arm’s length. & Customs houses, bonded warehouses, privateers’ brokers. & A “storm clause” traps a whole convoy in bonded storage; slip the goods out legally—or cleverly. \\
\hline
\textbf{The Temple Districts} (Sealing sites) & Public \emph{Brotherhood of the Seal} (ritual, civic charity) and hidden \emph{Custodians} (true-keeping, warding). & Jade pedestals, mnemonic chants, sealed vaults, desert hermitages. & A novice dies carrying a half-burned warning; follow the route before the desert swallows the rest. \\
\hline
\textbf{Kuvani Foothills} (North/east skirts) & Kuvani-descended riders under Dhaharan law; truce-markets and feud seasons. & Steppe ponies, horn bows, caravan outriders. & Feud-season spills onto the Silk road; negotiate a truce so the season’s last convoy can pass. \\
\hline
\end{tabular}
\end{center}

\paragraph{Faith and Festivals.}
Sky-feasts mark the first clear night after the rains; banners are judged for truth and craft. Breaking a sworn banner-oath is a public crime \emph{and} a sacrilege. Caravan gods are honored at waystones; at ports, Tempest priests bless departures while Dhahar’s kola-ash is thrown to the wind.

\paragraph{Law on the Road.}
Waystones guarantee right-of-passage to anointed caravans: harming them invokes both clan reprisals and march law. Quay charters cannot void waystone protections inland—merchants hire witnesses who know the difference.

\paragraph{Orders of the Seal.}
The public Brotherhood keeps libraries, schools, and civic rites around old foundations; the hidden Custodians maintain wards, train “rememberers,” and quietly move relics when politics lean dangerous. Both claim descent from the same vow; neither admits the other is wrong.

\paragraph{Trade Patterns.}
Caravans move pepper, resin, cotton, and carved jade north to river deltas and west to the Amrathian; iron, coin-credit, and timber flow back. In lean monsoons, resin and pilots fetch double; in fat ones, horses and guards do.

\subsection*{Nihon — “Isles of Quiet Steel”}
\textbf{Tagline:} Mountain isles, sea-lords, and sword schools that cut the wind.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Volcanic spines and cedar coasts; domains sworn to sea-lords and shrine courts.
  \item Duels of lineage styles; smiths fold storm and silence into steel.
  \item Exports: tempered blades, lacquer, pearl, dried fish. Imports: iron ore, horses, rare texts.
  \item Languages: Nihonese; shrine-cant among priestesses; Common in treaty ports.
  \item Hooks: A school’s secret kata is stolen before midsummer bouts; a reef-chart goes missing as typhoon bells ring.
\end{itemize}

\subsection*{Fharan Lands — “Sands of Frankincense and Fire”}
\textbf{Tagline:} Desert caravans, incense kingdoms, and star-watching citadels.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Oases string trade across red dunes; coastal incense cities fund desert fortresses.
  \item Tribal confederacies swear to city imams and salt princes; hospitality sacred, vendetta older still.
  \item Exports: frankincense, myrrh, glass, astrolabes. Imports: grain, timber, trained engineers.
  \item Languages: Fharan tongues; calculator cants in observatories; Common in caravanserais.
  \item Hooks: A sacred well is poisoned on the eve of treaty; a glassworks seeks protection from sand-cults.
\end{itemize}

\subsection*{Ashaan — “Thrones Under Sand”}
\textbf{Tagline:} River-lords, sun-temples, and tombs that still pay taxes.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Black silt kingdoms along a long river; green flood-bands bracketed by stone and desert.
  \item Viziers balance temple coffers and irrigation; desert cavalry patrol necropolis roads.
  \item Exports: grain, linen, natron, gold leaf, papyrus. Imports: timber, tin, mercenaries.
  \item Languages: Old Ashaani (liturgical), Bazaar Ashan (trade); Common with foreigners.
  \item Hooks: A tomb-audit goes lethal as ledgers vanish; a floodgate curse locks the river before planting.
\end{itemize}

\subsection*{Ayohkhan — “Emerald Archipelago of Smoke and Spice”}
\textbf{Tagline:} River-delta kingdoms, jungle thrones, and trade winds that choose their own kings.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Mangrove deltas and mountain jungles; stilt cities rule estuaries while hill citadels tax passes.
  \item Spice unions sway thrones; monsoon fleets leapfrog isles with pilot-priests.
  \item Exports: cinnamon, clove, camphor, bird-nest delicacies. Imports: iron, coin, scribes.
  \item Languages: Ayohkhan speech families; pilot litanies at sea; Common in entrepôts.
  \item Hooks: A monsoon calendar is forged to rig spice futures; hill rebels seize the only pass before harvest.
\end{itemize}

\chapter{The Gods and Powers}
Religion is both faith and faction. Each power gives rise to cults, orders, and schisms:
\begin{itemize}
  \item \textbf{Everflame} — Adar, Odur, Akilesh as facets of one holy fire. Seat of the Church in Ecktoria.  
  \item \textbf{Light} — Evangelical offshoot of the Everflame, rooted in Viterra.  
  \item \textbf{Khemesh \& Ráeyn} — Lord of Krakens and Lady of Storms, twin patrons of Kahfagia.  
  \item \textbf{Sky-Spirit of the Ykrul} — unnamed, shamanic, tied to omens.  
  \item \textbf{Dhahar} — war-god of the southern peninsula that bears his name.  
  \item \textbf{Ikasha, She Who Sleeps Within the Shadow} — Tulkani matron, whispered more than preached.  
  \item \textbf{Old Ancestors} — Dwarven reverence of kin, blood, and the Stone.  
\end{itemize}

\chapter{Cultures at a Glance}

\section{Humans}
Adaptable, ambitious, and diverse. Found across every province and sea.  
Affinity: \emph{Versatility}.  

\subsection{Sub-Cultures}
\subsubsection{Viterrans}
Highland pragmatists and fen traders. Known for their \emph{Knights of the Dawn} and a culture of logistics.  
XP Examples: Enhance Self (stamina, logistics), Acquire Assets (ferry rights, keeps), Learn Talents (road-law, dawn-muster).  

\subsubsection{Vhasians}
Old nobility and fractious dukes. Value bloodlines, alliances, and grandeur.  
XP Examples: Enhance Self (Presence), Acquire Assets (estates, alliances), Learn Talents (courtly intrigue).  

\subsubsection{Ecktorians}
Imperial remnants, gladiatorial traditions, and Church of the Flame zeal.  
XP Examples: Enhance Self (Body/Presence), Acquire Assets (arena, church patronage), Learn Talents (oratory, zealotry).  

\subsubsection{Acasians}
Petty warlords and bravos. Known for Silkstrand and mercenary companies.  
XP Examples: Enhance Self (blade skill), Acquire Assets (bravo gangs, ships), Learn Talents (dueling flourishes).  

\subsubsection{Ubral}
Clannish highlanders and lowlanders; guerrilla fighters and dwarven allies.  
XP Examples: Enhance Self (Stealth, guerilla tactics), Acquire Assets (clanholds, blood oaths), Learn Talents (hit-and-run mastery).  

\subsubsection{Tulkani}
Wanderers, performers, and traders across the north and Amrathian basin.  
XP Examples: Enhance Self (performance, cunning), Acquire Assets (wagon caravans, hidden networks), Learn Talents (witchcraft, shadowbinding).  

---

\section{Dwarves (Aeler)}
Stone-founded, guild-bound, collectivist.  
Affinity: \emph{Stone-Sense}.  

\subsection{Sub-Types}
\subsubsection{Mountain Dwarves}
Deep kingdom dwellers; focused on giants, dragons, and underrealms. Prestige: Forge-Patriarchs.  

\subsubsection{Hill Dwarves}
Surface allies of humans; fight Ykrul and trolls. Prestige: Spirit Shield Warriors.  

\subsubsection{Edgewalkers}
Outcast adventurers, risk-takers, and opportunists abroad. 

\subsubsection{True Masons}
Itenerate stone-clerics, craftsmen who maintain Aeler infrastructure abroad.

\subsubsection{Spirit Shields}
Traditionalist, spiritual warriors who uphold the old ways.

---

\section{Elves (Lethai)}
Elves see themselves as the first people, divided by philosophy.  

\subsection{Sub-Types}
\subsubsection{Wood Elves (Lethai-al)}
“Mist People,” tribal and fey; cycle-focused. Prestige: Nature Rangers, Backlash Soothers.  

\subsubsection{High Elves (Lethai-thora)}
Sequestered exiles of the mind; memory and continuity. Prestige: Echo-Walkers.  

\subsubsection{Grass / Dusk / Sundered Elves}
Nomadic high elves estranged from both cousins.  

---

\section{Ykrul}
Pragmatic warbands of horse, boar, and warg.  
Affinity: \emph{Blood Memory}.  

\subsection{Sub-Types}
\subsubsection{Steppe Ykrul}
Pastoral nomads of the Violet Steppes.  

\subsubsection{Mountain Ykrul}
Stockier, fierce, more reclusive clans of the Aelerian highlands.  

\subsubsection{River Ykrul}
Wargboat raiders and coastal plunderers. Prestige: Warglords.  

---

\section{Gnomes (Aelinnel)}
Affinity: \emph{Numerical Insight}. Known for incomprehensible mathematics and craft.  

\subsection{Sub-Types}
\subsubsection{Rock Gnomes}
Burrow-dwellers of Wrentfell; gemstone engineers.  

\subsubsection{Forest Gnomes}
Valewood dwellers; more attuned to nature and fey.  

\subsubsection{Faetae (Winged Gnomes)}
Rare, fragile, and fae-blooded with insectile wings. Can fly but this often leads to complications.  

---

\section{Halflings (Aelaeram)}
Affinity: \emph{Hearth Sense}. Pastoral and resilient.  

\subsection{Sub-Types}
\subsubsection{Hill Halflings}
The most common; Amedell farmers and traders.  

\subsubsection{River Halflings}
Bargefolk and ferry-keepers along the Belworth.  

\subsubsection{Wander Halflings}
Caravaners and adventurers who leave the hearth to find fortune.   

\chapter{The XP Paths in Practice}
\section{Enhance Self}
Investing in raw power and skill.  
\begin{quote}
A Viterran knight spends XP to raise Body and Melee, becoming a duelist feared on the field.  
\end{quote}

\section{Acquire Assets}
Investing in the world around you.  
\begin{quote}
A Vhasian Duke invests XP into a Keep (off-screen) and house guards (on-screen).  
\end{quote}

\section{Learn Talents}
Investing in uniqueness.  
\begin{quote}
A Tulkani shadowbinder learns to weave concealment magic, turning night into an ally.  
\end{quote}

\chapter{Putting It Together}
Every player’s path blends the three. One may stand alone as a Solo, another weave allies as a Mastermind, another balance the two.  
Fate’s Edge rewards the choice, not just the outcome.  
The world reacts to where you put your weight—on the self, the world, or the myth you embody.

\chapter{Stock NPCs}

Not every companion, pet, or hireling needs to be custom-built.  
This section provides quick archetypes for \textbf{players} (on-screen allies) and \textbf{GMs} (encounters, foils, and rivals).

\section{Player Companions}
On-screen followers are bought with XP, and their abilities are measured by a \emph{Skill Cap}. They support but also create vulnerabilities.  

\subsection{Familiars \& Pets}
Bound by bond or training. Useful for scouting, flavor, and niche aid.  
\begin{description}[leftmargin=2cm]
  \item[Shadow-Cat] Cap 2 in Stealth. Can climb, slip through cracks, and carry small items. Backlash: draws bad omens.  
  \item[Crow Messenger] Cap 2 in Perception. Can deliver notes or warn of danger. Backlash: nosy, sometimes lies.  
  \item[Hound of the Fens] Cap 3 in Tracking. Keen nose, loyal defender. Backlash: loud bark alerts enemies.  
\end{description}

\subsection{Hirelings \& Followers}
Professionals who lend their skill to a player’s action.  
\begin{description}[leftmargin=2cm]
  \item[Bodyguard] Cap 4 in Melee. Grants bonus dice in combat, but draws fire.  
  \item[Scribe] Cap 3 in Lore. Keeps records, interprets contracts, whispers legal loopholes.  
  \item[Scout] Cap 3 in Survival. Knows paths, can extend a party’s range, but prone to wanderlust.  
\end{description}

\section{GM Stock NPCs}
These NPCs are not built with XP, but with intent. Each carries a role in pacing and tension.  

\subsection{Encounters}
\begin{description}[leftmargin=2cm]
  \item[Bandit Skirmisher] Body 2, Wits 2, Skills: Melee 2, Stealth 1. Light armor, opportunistic, breaks if pressed.  
  \item[Ykrul Rider] Body 4, Wits 3, Skills: Riding 3, Melee 3. Mobile, brutal, fights for plunder.  
  \item[Street Bravo] Presence 3, Body 2, Skills: Dueling 3. Quick to anger, duels to prove honor.  
\end{description}

\subsection{Foils \& Rivals}
\begin{description}[leftmargin=2cm]
  \item[Ambitious Scribe] Wits 3, Presence 3, Skills: Intrigue 3, Lore 2. Always knows a rumor, but hides motives.  
  \item[Mercenary Captain] Body 4, Spirit 3, Skills: Command 3, Melee 4. Pragmatic, can be ally or threat.  
  \item[Flame Preacher] Presence 4, Spirit 3, Skills: Oratory 4, Faith 3. Incites mobs, can spark holy trouble.  
\end{description}

\chapter{Gameplay Demo}

\section*{At the Table}
To illustrate the flow of play in \textit{Fate’s Edge}, here is a short example.  
The players are exploring an abandoned watchtower in the Mistlands.  

\subsection*{Scene: Climbing the Tower}

\textbf{GM:} The wooden stairs are rotted through. If you want to reach the beacon chamber, you’ll need to climb the stone wall.  

\textbf{Player (Rin, the Scout):} I’ll scramble up using my dagger for handholds. That’s Body + Athletics, right?  

\textbf{GM:} Yes. What’s your approach?  

\textbf{Rin:} Intricate — I describe how I wedge the dagger into the cracks, kicking off with precision.  

\textbf{GM:} Great. Roll 4 dice.  

\textit{The player rolls 6, 5, 1, 1. Two successes, two Complication Points. Because it’s Intricate, Rin rerolls both 1s: 7 and 2. Now three successes total.}  

\textbf{GM:} You make it to the top, but I’ll spend those two Complication Points.  
\textit{(Draws from the Deck of Consequences: Clubs 3 = “Equipment Strain.”)}  
Your dagger is stuck in the stone, bent and useless.  

\textbf{Rin:} Worth it — I pull myself onto the platform.  

---

\subsection*{Scene: The Beacon Chamber}

\textbf{GM:} The chamber reeks of smoke and salt. A hunched figure stirs the ashes — a Mistborn wraith.  

\textbf{Player (Kael, the Warlock):} I’ll cast a firebolt to drive it back. Channel first: Wits + Arcana.  

\textit{Kael rolls 5 dice: 8, 6, 3, 1, 1. Two successes, two Complication Points.}  

\textbf{Kael:} Detailed approach — I reroll one 1. That’s a 7. Three successes.  

\textbf{GM:} You’ve built strong Potential. Now Weave it with Fire.  

\textit{Kael rolls Wits + Fire Art: 6, 5, 2, 1. One success, one more Complication Point.}  

\textbf{GM:} The firebolt streaks out — the wraith hisses and withdraws. But I’ve got three Complication Points total.  
\textit{(Draws from Deck: Hearts 2 = “Emotional Cost.”)}  
The firebolt sparks panic in your companions — the Mistlands mist churns, and the shadows deepen. Everyone takes a level of Fatigue from the oppressive dread.  

---

\subsection*{Takeaway}
This example shows the rhythm of \textit{Fate’s Edge}:
\begin{itemize}
  \item Players describe their approach, choosing Basic, Detailed, or Intricate.  
  \item Dice are rolled; successes move the story forward, but Complication Points fuel the GM.  
  \item The GM uses Complications (or the Deck of Consequences) to keep tension alive.  
  \item Every roll advances the fiction, never just “succeed/fail.”  
\end{itemize}

\begin{tcolorbox}[title=Player Tip: How to Earn Boons,colback=black!2,colframe=black!40!white]
  Boons are not random handouts—they reward when you take \textbf{real risks} and \textbf{move the story forward}.  
  
  \begin{itemize}
    \item Take bold actions, even if they might fail.  
    \item Describe what you do in the fiction—make it vivid.  
    \item Accept complications and setbacks; they fuel your future.  
    \item Don’t “fish” with meaningless rolls—no story, no Boon.  
  \end{itemize}
  
  \textbf{Remember:} A Boon is proof that you dared, failed, and grew.
  \end{tcolorbox}

\section*{Min–Max Showcases (Player Options)}

\subsection*{The Unstoppable Blade (Solo Peak)}
\textbf{Concept:} A duelist honed to a lethal edge; controls any fight that involves their blade.

\paragraph{Dice Pools}
\begin{itemize}
  \item \textbf{Strike/Parry:} Body 4 + Melee 3 = \textbf{7d10}
  \item \textbf{Read Foe / Initiative:} Wits 3 + (no skill) = \textbf{3d10}
  \item \textbf{Social Baseline:} Presence 2 + (no skill) = \textbf{2d10}
\end{itemize}

\paragraph{XP Build (Target 32; rules-legal with deficit trades)}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Category} & \textbf{Pick} & \textbf{XP} \\
\midrule
Attributes & Body 4 & 12 \\
           & Wits 3 & 9 \\
Skills     & Melee 3 & 12 \\
Talents    & \emph{Battle Instincts} & 4 \\
\midrule
\multicolumn{2}{r}{\textit{Subtotal}} & \textbf{37} \\
\multicolumn{2}{r}{Starting Complications $\rightarrow$ XP (3 picks)} & \textbf{-3} \\
\multicolumn{2}{r}{Boon Bonds (4 Boons $\Rightarrow$ 2 XP)} & \textbf{-2} \\
\midrule
\multicolumn{2}{r}{\textit{Net Start}} & \textbf{32} \\
\bottomrule
\end{tabular}

\paragraph{Suggested Starting Complications (choose 3)}
\emph{Cursed Token}; \emph{Notoriety (dueling circles)}; \emph{Blood Debt}.  
\textbf{Boon Bonds:} Tie to two other PCs (mutual) for 4 Boons (2 XP).

\paragraph{Table Notes}
Always use \textbf{Intricate} descriptions in combat to re-roll 1s and add flourish. Terrifying in a fight, intentionally underpowered outside it — rely on allies for social/puzzle scenes.

\bigskip

\subsection*{The Ghost in the System (Mastermind Peak)}
\textbf{Concept:} A fixer who solves problems off-screen with networks, safehouses, and favors.

\paragraph{Dice Pools}
\begin{itemize}
  \item \textbf{Deception/Setup:} Presence 3 + Subterfuge 2 = \textbf{5d10}
  \item \textbf{Read Angles:} Wits 3 + Insight 2 = \textbf{5d10}
  \item \textbf{On-Screen Muscle:} (low; bring a bodyguard follower if desired)
\end{itemize}

\paragraph{XP Build (Target 32; rules-legal)}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Category} & \textbf{Pick} & \textbf{XP} \\
\midrule
Attributes & Presence 3 & 9 \\
           & Wits 3 & 9 \\
Skills     & Subterfuge 2 & 6 \\
           & Insight 2 & 6 \\
Off-Screen Asset & \textbf{Standard Safehouse Network} & 8 \\
\midrule
\multicolumn{2}{r}{\textit{Subtotal}} & \textbf{38} \\
\multicolumn{2}{r}{Starting Complications $\rightarrow$ XP (4 picks)} & \textbf{-4} \\
\multicolumn{2}{r}{Boon Bonds (4 Boons $\Rightarrow$ 2 XP)} & \textbf{-2} \\
\midrule
\multicolumn{2}{r}{\textit{Net Start}} & \textbf{32} \\
\bottomrule
\end{tabular}

\paragraph{Suggested Starting Complications (choose 4)}
\emph{Debt Mark}; \emph{Fractured Loyalty}; \emph{Blackmail}; \emph{Owed Favor}.  
\textbf{Boon Bonds:} Tie to two PCs for 4 Boons (2 XP).  
\textbf{Activation Fuel:} Use \textbf{1 Boon or 2 XP} to activate the Safehouse Network for a scene-level ask.

\paragraph{Table Notes}
You “win” by \emph{not rolling}. Spend Boons to open doors, farm Boons with controlled risks. Fragile if the Boon engine dries up.

\bigskip

\subsection*{The Voice of God (Face Peak)}
\textbf{Concept:} A presence that ends fights with words, ritual, and authority.

\paragraph{Dice Pools}
\begin{itemize}
  \item \textbf{Persuade/Compel:} Presence 3 + Diplomacy 3 + \emph{Silver Tongue} = \textbf{7d10}
  \item \textbf{Rally/Direct:} Presence 3 + Command 1 (+ talent if framed) = \textbf{4–5d10}
  \item \textbf{Resolve/Resist:} Spirit 3 = \textbf{3d10}
\end{itemize}

\paragraph{XP Build (Target 32; rules-legal with deficit trades)}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Category} & \textbf{Pick} & \textbf{XP} \\
\midrule
Attributes & Presence 3 & 9 \\
           & Spirit 3 & 9 \\
Skills     & Diplomacy 3 & 12 \\
           & Command 1 & 2 \\
Talents    & \emph{Silver Tongue} & 3 \\
\midrule
\multicolumn{2}{r}{\textit{Subtotal}} & \textbf{35} \\
\multicolumn{2}{r}{Starting Complications $\rightarrow$ XP (2 picks)} & \textbf{-2} \\
\multicolumn{2}{r}{Boon Bonds (2 Boons $\Rightarrow$ 1 XP)} & \textbf{-1} \\
\midrule
\multicolumn{2}{r}{\textit{Net Start}} & \textbf{32} \\
\bottomrule
\end{tabular}

\paragraph{Suggested Starting Complications (choose 2)}
\emph{Notoriety (public figure)}; \emph{Enemy at Large}.  
\textbf{Boon Bonds:} Tie to one PC for 2 Boons (1 XP), or two PCs for 4 Boons (2 XP) and bank 1 XP or add Insight~1.

\paragraph{Table Notes}
\emph{Intricate} speeches reroll 1s and add social flourishes (“he becomes a quiet informant”). You fold scenes without steel, but zealots/monsters resist your leverage. Keep allies between you and blades.

\bigskip

\paragraph{GM Sidebar (All Three)}
\begin{itemize}
  \item Reward \textbf{Intricate} declarations with rerolls and world-reactive flourishes.
  \item Pressure each archetype at its weakest axis (Blade: politics; Ghost: resource drought; Voice: brute force) while letting them shine often.
  \item Enforce Off-Screen Asset activation: \textbf{1 Boon or 2 XP} per concrete ask; if refused, show friction instead of hard no.
\end{itemize}

\end{document}
