\section{Foundation Talents (2-4 XP)}

These talents are available to any Runic Warrior regardless of Patron. They represent the basic physical disciplines, resilience, and practical skills that define the warrior's path.

\subsection{Flesh Codex (2 XP)}

\textbf{Requirement:} Body 3+

Your Rites are engraved on your body—tattoos, brands, or scarification. You have no physical Codex to steal, confiscate, or destroy.

\textbf{Effect:}
\begin{itemize}
\item Your Codex cannot be stolen or confiscated.
\item You always have access to your Rites (no action required to retrieve a Codex).
\item Damage to your skin (burns, deep wounds) may damage your Codex at the GM's discretion.
\item In cultures that respect flesh-writing, you gain +1 die to social rolls; in cultures that fear it, you suffer -1 die.
\end{itemize}

\subsection{Living Thiasos (3 XP)}

\textbf{Requirement:} A living Thiasos (warhorse, hound, wolf-ancestor, etc.)

Your Thiasos is not a pet. It is a partner.

\textbf{Effect:}
\begin{itemize}
\item Your Thiasos gains +1 Cap (maximum 5).
\item Once per scene, your Thiasos may act independently without requiring your action.
\item Your Thiasos is more resilient: when it would be reduced to 0 Harm, it remains conscious but Winded (-1 die to all actions) until healed.
\item You may sense your Thiasos's emotional state (calm, alert, frightened, hostile) as long as it is within Far range.
\end{itemize}

\subsection{Oath-Bound (4 XP)}

\textbf{Requirement:} You have sworn a formal oath before a witness.

Your word is your bond. Your bond is your strength.

\textbf{Effect:}
\begin{itemize}
\item Once per session, when you act in direct service of a sworn oath, you may reroll a failed Body roll.
\item If you break your sworn oath, you lose access to this talent until you perform a witnessed atonement.
\item The oath must be specific (e.g., "I will protect this village," not "I will be good"). Vague oaths carry no power.
\end{itemize}

\subsection{Scarred Resilience (3 XP)}

\textbf{Requirement:} Body 2+

You have learned to read your own pain.

\textbf{Effect:}
\begin{itemize}
\item Once per scene, you may ignore 1 Fatigue from physical exertion (running, climbing, fighting, or invoking a Rite that costs Fatigue).
\item You may use this talent after the Fatigue is marked.
\item You cannot use this talent to ignore Fatigue from Obligation or magical backlash.
\end{itemize}

\subsection{Talking Weapon (2 XP)}

\textbf{Requirement:} A sentient Thiasos (weapon, shield, or object)

Your sentient Thiasos can speak aloud (not just in your mind). It has its own personality, opinions, and occasionally inappropriate commentary.

\textbf{Effect:}
\begin{itemize}
\item Your Thiasos may serve as a witness for Rites that require one (it speaks, therefore it testifies).
\item The GM may use your Thiasos to deliver hints, warnings, or comic relief.
\item Your Thiasos has its own agenda. It may refuse to cooperate if it disagrees with your actions.
\item Once per session, you may ask your Thiasos one question about a creature, object, or location it has witnessed. The GM must answer truthfully, though the answer may be cryptic.
\end{itemize}

\subsection{Weapon Mastery (4 XP)}

\textbf{Requirement:} Body 2+

Choose a weapon category (swords, axes, spears, hammers, unarmed, or thrown). You have trained extensively with this weapon type.

\textbf{Effect:}
\begin{itemize}
\item When using a weapon from your chosen category, you gain +1 die to attack rolls.
\item You may reroll one 1 on damage with your chosen weapon.
\item You may take this talent multiple times for different weapon categories.
\end{itemize}

\subsection{Shield Bearer (4 XP)}

\textbf{Requirement:} Body 2+, a shield

You have trained to use a shield not as a passive defense but as an active tool of combat.

\textbf{Effect:}
\begin{itemize}
\item When wielding a shield, you gain +1 Armor.
\item You may use the Protect action to interpose yourself between an ally and an attack, taking the Harm yourself.
\item When you take the Defend action, adjacent allies also gain the benefit of your shield (+1 Armor) until your next turn.
\end{itemize}

\subsection{Second Wind (2 XP)}

\textbf{Requirement:} Body 2+

You know how to push through exhaustion when others would fall.

\textbf{Effect:}
\begin{itemize}
\item Once per session, as an action, you may clear 2 Fatigue and remove one minor Condition (Dazed, Winded, Shaken).
\item You cannot use this talent while at 0 Harm (incapacitated).
\item After using Second Wind, you mark 1 Obligation—the debt of borrowed endurance.
\end{itemize}

\subsection{Iron Stomach (3 XP)}

\textbf{Requirement:} Body 2+

You have eaten worse things than whatever is in front of you.

\textbf{Effect:}
\begin{itemize}
\item You are immune to mundane poisons and spoiled food.
\item You suffer -1 die from complications caused by ingested toxins or diseases.
\item You may consume raw meat, rotten provisions, or questionable fungi without penalty.
\end{itemize}

\subsection{Keen Senses (2 XP)}

\textbf{Requirement:} Wits 2+

You have trained your senses to notice what others overlook.

\textbf{Effect:}
\begin{itemize}
\item +1 die to Perception rolls.
\item You cannot be surprised in natural terrain (forest, plains, mountains, swamps) unless the threat is magical.
\item You may identify the general direction and distance of a sound within Far range.
\end{itemize}

\subsection{Disciplined Body (2 XP)}

\textbf{Requirement:} Body 2+

You have trained your body to endure what would break others.

\textbf{Effect:}
\begin{itemize}
\item You may ignore 1 point of Fatigue from physical exertion once per scene.
\item You may hold your breath for twice as long as normal (test Body + Athletics vs DV 3 to extend to 1 minute per success).
\item You suffer -1 die from environmental penalties caused by extreme temperatures, thin air, or difficult terrain.
\end{itemize}

\subsection{Talent Synergies}

Some Foundation Talents work well together:

\begin{itemize}
\item \textbf{Flesh Codex + Scarred Resilience:} Your body is both Codex and shield. You cannot be disarmed, and you ignore the first Fatigue each scene.
\item \textbf{Living Thiasos + Talking Weapon:} Your Thiasos is a partner in truth—a witness that speaks, fights, and has opinions. The GM may use it as a narrative device.
\item \textbf{Weapon Mastery + Shield Bearer:} You are a warrior's warrior—deadly with your chosen weapon and a wall of protection for your allies.
\item \textbf{Second Wind + Disciplined Body:} You recover from exhaustion faster than anyone expects. The debt of borrowed endurance is worth the price.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
