\section{Advanced Talents (6-8 XP)}

These talents represent the refined skills and deeper bonds of a seasoned Runic Warrior. They require significant training, battlefield experience, or a particularly strong covenant with one's Patron.

\subsection{Engraved Defense (6 XP)}

\textbf{Requirement:} Your Codex is a physical object (gorget, shield, staff, or sentient blade)

Your Codex is not just a record. It is armor. The Rites carved into its surface deflect blows as surely as steel.

\textbf{Effect:}
\begin{itemize}
\item Your Codex grants +1 Armor.
\item When you would take Harm, you may instead mark 1 Obligation and reduce the Harm by 1 (to a minimum of 0).
\item You cannot use this talent if your Codex is Compromised or destroyed.
\item The Obligation marked this way is visible—a new crack in the shield, a fresh scratch on the gorget, a dulling of the sentient blade's edge.
\end{itemize}

\subsection{Shared Burden (6 XP)}

\textbf{Requirement:} A living Thiasos

You and your Thiasos share more than a covenant. You share weight. The bond between you is strong enough to transfer exhaustion.

\textbf{Effect:}
\begin{itemize}
\item When you mark Fatigue, you may transfer 1 point of it to your Thiasos.
\item Your Thiasos becomes Winded (-1 die to all actions) for the rest of the scene.
\item You may use this talent once per scene.
\item If your Thiasos is already Winded, you cannot transfer additional Fatigue to it.
\end{itemize}

\subsection{Witnessed Scar (8 XP)}

\textbf{Requirement:} You have at least three permanent Obligation scars

Each scar is a payment made. Each payment is a memory. Your scars do not just record—they testify.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may touch a scar and recall the event that caused it in perfect detail.
\item Gain +1 die on any roll directly tied to that event (tracking a foe you fought before, recalling a Patron's demand, navigating a place you bled in).
\item If you touch a scar while invoking a Rite, the Rite's Obligation cost is reduced by 1 (minimum 0). The scar remembers the debt—and accepts the payment.
\end{itemize}

\subsection{Iron Resolve (7 XP)}

\textbf{Requirement:} Body 3+, Disciplined Body or similar resilience talent

You do not fall easily. When others would collapse, you plant your feet and refuse.

\textbf{Effect:}
\begin{itemize}
\item When you would be reduced to 0 Harm, you may instead remain standing with 1 Harm and mark 2 Fatigue.
\item You may use this talent once per scene.
\item The Fatigue marked this way cannot be transferred to your Thiasos or ignored by Scarred Resilience. The debt is yours alone.
\end{itemize}

\subsection{Flowing Strike (6 XP)}

\textbf{Requirement:} Weapon Mastery (any), Body 3+

Your attacks flow into movement. You do not strike and stand still—you strike and flow.

\textbf{Effect:}
\begin{itemize}
\item When you successfully hit a target, you may immediately make a free repositioning move (up to Near distance) without provoking opportunity attacks.
\item You may use this ability once per scene.
\item If you are using a weapon from your chosen Weapon Mastery category, you may use this ability twice per scene.
\end{itemize}

\subsection{Flurry (6 XP)}

\textbf{Requirement:} Weapon Mastery (any), Body 3+

You attack with blinding speed, striking twice before your opponent can react.

\textbf{Effect:}
\begin{itemize}
\item Once per scene, you may make two attacks against the same target as a single action. Resolve each attack separately.
\item After using Flurry, you suffer -1 die on your next action (the exertion leaves you slightly off-balance).
\item If both attacks hit, you may immediately make a free repositioning move (as Flowing Strike).
\end{itemize}

\subsection{Battlefield Awareness (6 XP)}

\textbf{Requirement:} Wits 3+

You have learned to read the chaos of combat. The battlefield speaks—you have learned to listen.

\textbf{Effect:}
\begin{itemize}
\item At the start of any combat scene, you may ask the GM one tactical question: "Which enemy is most dangerous?", "Where is the best cover?", "What is the escape route?" The GM must answer truthfully.
\item You cannot be surprised in combat unless the ambush is magical or the attackers are invisible.
\item You gain +1 die to detect hidden enemies or traps during combat.
\end{itemize}

\subsection{The Blade Remembers (6 XP)}

\textbf{Requirement:} A sentient Thiasos (weapon)

Your sentient Thiasos has witnessed every battle you have fought. It remembers the enemies you have faced, the strikes that landed, the blows that nearly killed you.

\textbf{Effect:}
\begin{itemize}
\item Once per session, your Thiasos may recall a past enemy's fighting style. Gain +1 die to defend against a similar foe (same species, same weapon type, same tactics).
\item Your Thiasos may warn you of an incoming attack from an enemy it has seen before. The first attack against you in a scene from a familiar enemy suffers -1 die.
\item Your Thiasos may testify to your deeds. In any court that respects sentient witness, its word carries weight.
\end{itemize}

\subsection{The Hilt Listens (6 XP)}

\textbf{Requirement:} A sentient Thiasos (weapon)

Your sentient Thiasos can hear conversations within Near range even when you are not present—if you leave it in a room, hidden in a sheath, or placed on a shelf.

\textbf{Effect:}
\begin{itemize}
\item When you leave your Thiasos in a location (as a free action), it may listen to any conversation within Near range.
\item At the end of the scene, your Thiasos reports what it heard. It does not interpret—it repeats.
\item Your Thiasos can identify voices it has heard before, even if the speaker is disguised or altered.
\item Your Thiasos cannot move on its own. It can only listen where it is left.
\end{itemize}

\subsection{Tide-Step (6 XP)}

\textbf{Requirement:} Raéyn or The Traveler, Body 3+

You move with the rhythm of water or the turn of the road. The world flows beneath your feet.

\textbf{Effect:}
\begin{itemize}
\item Once per session as a free action, you may treat any water-based or road-based difficult terrain as stable ground for one exchange.
\item You may move across the surface of water as if it were solid for up to Near distance.
\item Mark 1 Fatigue after using this talent. The tide does not lift without cost.
\end{itemize}

\subsection{Storm Caller (6 XP)}

\textbf{Requirement:} Xhak'Thul, Spirit 3+

The storm speaks through you. Your voice is thunder; your presence is lightning.

\textbf{Effect:}
\begin{itemize}
\item Once per scene, as a free action, you may cause a peal of thunder to roll across the zone.
\item All enemies within Near must test Resolve (DV 3) or lose their next minor action (distracted by the sudden crash).
\item If you are in an active storm (rain, wind, thunder), the Resolve test is DV 4.
\end{itemize}

\subsection{Shadow Meld (6 XP)}

\textbf{Requirement:} Ikasha, Wits 3+

You step into shadow and emerge from shadow. The dark is your ally—and your highway.

\textbf{Effect:}
\begin{itemize}
\item Once per session as an action, step into any shadow larger than yourself and emerge from another shadow within Near range.
\item You cannot be targeted while moving between shadows.
\item If you emerge from a shadow in a Dominant Position, your next attack gains +1 Effect.
\item Mark 1 Fatigue after using this talent.
\end{itemize}

\subsection{Talent Synergies}

Advanced Talents combine with Foundation Talents for powerful effects:

\begin{itemize}
\item \textbf{Engraved Defense + Flesh Codex:} Your body is your Codex; your scars are your armor. You cannot be disarmed, and your defenses are always present.
\item \textbf{Shared Burden + Living Thiasos:} You and your Thiasos endure together. Neither falls alone. You may transfer Fatigue to your partner, and it fights on despite wounds.
\item \textbf{Witnessed Scar + Flesh Codex:} Your scars are a public ledger. You may touch a scar to recall a memory—or to reduce a Rite's cost. The debt is visible; the payment is final.
\item \textbf{Iron Resolve + Scarred Resilience:} You ignore the first Fatigue each scene, and when you would fall, you stand. You are nearly unkillable—until the debt comes due.
\item \textbf{Flowing Strike + Flurry:} Attack twice, then reposition. You are a whirlwind of steel and movement. Enemies cannot pin you down.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
