\section{Prestige Talents (10-12 XP)}

These talents represent the pinnacle of a Runic Warrior's path—warriors who have transcended ordinary limits, whose scars tell stories of a hundred battles, whose bond with their Patron is so deep it has reshaped their very flesh. These talents are not learned lightly, and they are never forgotten.

\subsection{The Unbroken Vow (10 XP)}

\textbf{Requirement:} You have kept a sworn oath for at least one year (in-game time)

Some vows are not chains. They are bridges. You have kept your word so long that the keeping has become part of you.

\textbf{Effect:}
\begin{itemize}
\item When you swear an oath before a witness, you gain +2 dice to all actions that serve that oath.
\item Breaking the oath causes you to immediately mark 2 Harm (Stress) and lose access to this talent until you perform a witnessed atonement.
\item You may have only one Unbroken Vow active at a time. The oath must be specific and measurable (e.g., "I will protect this village until the spring thaw," not "I will be good").
\item When you fulfill the oath, you may immediately clear 2 Fatigue and remove one Condition. The bridge holds.
\end{itemize}

\subsection{Avatar of the Codex (12 XP)}

\textbf{Requirement:} Tier III+, your Codex is a physical object

You do not carry your Codex. You become it. The Rites engraved on your gorget, your shield, or your flesh are not merely records—they are you.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may become the living embodiment of your Codex for one scene.
\item Your Body increases by 2 (to a maximum of 6).
\item Your Thiasos fights beside you as a second combatant (taking its own actions).
\item Any Rite you invoke during this scene costs -1 Obligation (minimum 0).
\item At the end of the scene, you mark 2 Fatigue and 1 Obligation. The cost of becoming more than human is that you must remember you were human.
\end{itemize}

\subsection{The Scar That Remembers (11 XP)}

\textbf{Requirement:} Witnessed Scar, Tier III+

Your scars do not just record. They testify. They speak when you cannot. They remember what you wish you could forget.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may touch a scar and speak a single sentence about the event that caused it.
\item That sentence is considered witnessed truth. In a hedge-court, it carries the weight of sworn testimony. Before a Patron, it is a binding statement.
\item You cannot use this talent on the same scar twice. Each scar testifies once, then is spent—faded, silent, released.
\item If you use this talent to testify against someone who wronged you, you may also mark 1 Boon (the scar's vengeance).
\end{itemize}

\subsection{The Unburdened Crossing (12 XP)}

\textbf{Requirement:} The Pale Shepherd or Oath of Mercy and Grace, Spirit 4+

You have guided souls across the threshold. You have witnessed the moment when breath becomes silence, when weight becomes memory. You are not afraid of death—you are acquainted with it.

\textbf{Effect:}
\begin{itemize}
\item Once per session, when a willing creature (including yourself) is about to die, you may guide them across the threshold. They die peacefully, without fear.
\item Before they pass, they may speak one final truth. That truth is considered witnessed; it cannot be denied by any who hear it.
\item You gain a permanent memory of their last words, which may be used as a \textit{Truth Compass} (+1 die to any roll directly related to fulfilling their final wish).
\item You gain 1 permanent Corruption. The Shepherd does not guide for free.
\end{itemize}

\subsection{The Spiral's Final Witness (12 XP)}

\textbf{Requirement:} Oath of Mercy and Grace, Tier III+

You have learned that mercy is not weakness—it is a weapon. The spiral remembers every kindness, every spared life, every offered cup.

\textbf{Effect:}
\begin{itemize}
\item Once per session, when a creature with an active Promise Clock (from any of your Rites) would deal a killing blow to an ally, you may interpose yourself.
\item The blow is redirected to you, and you suffer the Harm instead.
\item The Promise Clock fills immediately, and the attacker suffers the consequence (Harm, Condition, or Curse as described in the Rite that created the clock).
\item You cannot use this talent again until you have spent a downtime scene in meditation or prayer.
\end{itemize}

\subsection{Apotheosis of Hunger (12 XP)}

\textbf{Requirement:} The Savage Heart or Malachai, Body 4+, Spirit 3+

You have tasted the raw hunger of the predator. You have felt the curse of the thorn. You are no longer simply a warrior—you are an appetite wearing skin.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may enter a state of perfect consumption for one scene.
\item You ignore all penalties from Fatigue.
\item Any time you would take Harm from a physical attack, you may instead convert the Harm into 1 Fatigue and gain +1 die on your next action (the hunger feeds on the blow).
\item When the scene ends, you suffer Harm 2 (Stress) as your body rejects the excess. You cannot use this talent again until you have consumed a significant meal (at least a full day's rations) in a ritual manner.
\end{itemize}

\subsection{Runebound Resilience (12 XP)}

\textbf{Requirement:} Tier III+, Codex, Spirit 4+

You have learned to transform the very essence of harm into deeper communion with your Patron. Your flesh is a ledger. Every wound is a prayer.

\textbf{Effect:}
\begin{itemize}
\item \textbf{Harm Conversion:} Once per round, you may convert up to 2 Harm levels to +2 Obligation instead of taking the damage.
\item \textbf{Obligation Endurance:} While at 6+ Obligation segments, gain +1 die to all Rite invocations.
\item \textbf{Patron's Shield:} When reduced to Harm 3, you may immediately invoke any known Rite without marking additional Obligation.
\item \textbf{Debt Fortification:} At 8+ Obligation segments, ignore the first Harm 1 each round.
\item \textbf{Once per scene limit} on the Harm Conversion ability.
\end{itemize}

\subsection{Patron's Favored (12 XP)}

\textbf{Requirement:} Tier III+, Codex, Familiar, 8+ total Obligation segments

Your Patron has marked you as specially favored, granting you unprecedented access to their power with reduced cost. You carry their blessing openly now, and their influence spreads like wildfire to those who would walk in your shadow.

\textbf{Effect:}
\begin{itemize}
\item \textbf{Low Rite Discount:} Reduce Obligation cost for Low Rites by 1 (minimum 0).
\item \textbf{Empower:} Once per scene, spend 1 Action + 1 Boon to empower a target (including yourself). The target gains +1 Corruption level per round for 3 rounds and +1 die to all rolls while empowered. Corruption lasts until the end of the scene.
\item \textbf{Empower Push:} Spend 1 Action + 1 Boon + 3 Obligation to empower a target (or targets) with +3 Corruption levels until end of scene, +1 die to all rolls, and the target gains Corruption fatigue at the end of the scene.
\item \textbf{Enhanced Gift:} Your Patron's Gift (Imbuement) now provides +2 dice instead of +1.
\item \textbf{Debt Bound:} Cannot Empower again until all Obligation is cleared.
\end{itemize}

\subsection{Talent Synergies}

Prestige Talents represent the culmination of a Runic Warrior's path:

\begin{itemize}
\item \textbf{The Unbroken Vow + Avatar of the Codex:} You become the living embodiment of your oath. Your Codex is your flesh; your vow is your armor. For one scene, you are nearly unstoppable—but the cost is high.
\item \textbf{The Scar That Remembers + Witnessed Scar:} Your scars are a public ledger. You may touch a scar to recall a memory, reduce a Rite's cost, or testify to the truth. Each scar speaks once, then is spent.
\item \textbf{The Unburdened Crossing + The Spiral's Final Witness:} You guide the dying and you shield the living. You are the threshold and the guardian. Mercy and death walk beside you.
\item \textbf{Apotheosis of Hunger + Runebound Resilience:} You are hunger and obligation made flesh. You convert harm to power, and power to hunger. The cycle is endless—until it consumes you.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
