\section{Patron-Specific Talents}

These talents are available only to Runic Warriors sworn to specific Patrons or Patron families. Each represents a unique expression of that Patron's power channeled through the warrior's body.

\subsection{The Vowed (Oaths, Law, and Judgment)}

\subsubsection{Dawn's Judgment (Oath of Flame and Light) — 6 XP}

\textbf{Requirement:} Oath of Flame and Light, Body 3+

When you invoke Rite of Dawn's Judgment, you may choose to extend the effect. Your blade remains kindled for the remainder of the scene, not just a single strike. Each additional attack beyond the first costs 1 Fatigue.

The flame does not tire. You might.

\subsubsection{Unquenchable (Oath of Flame and Light) — 4 XP}

\textbf{Requirement:} Oath of Flame and Light

Your flame cannot be extinguished by mundane means. Wind, water, and darkness do not affect your kindled blade or dawnfire Rites. Magical extinguishing requires a contested roll (Spirit + Lore vs. your Spirit + Presence).

The fire remembers who lit it.

\subsubsection{Spiral Counter (Oath of Mercy and Grace) — 6 XP}

\textbf{Requirement:} Oath of Mercy and Grace, Body 3+

When a creature with an active Promise Clock (from any of your Rites) attacks you, you may spend 1 Boon to advance that clock by 1 segment immediately, before the attack resolves.

The spiral tightens. The debt accelerates.

\subsubsection{The Unbuckled Gorget (Oath of Mercy and Grace) — 4 XP}

\textbf{Requirement:} Oath of Mercy and Grace

Once per session, you may remove a piece of armor (helm, gauntlet, or gorget) to show vulnerability. Gain +1 die to social rolls with honorable enemies, but suffer +1 Harm from the next attack against you.

The mercy is real. The risk is real. The armor opens.

\subsubsection{Cite Precedent (Mykkiel) — 4 XP}

\textbf{Requirement:} Mykkiel, Lore 2+

Once per session, when you invoke a Rite that establishes a judgment or binds a covenant, you may declare that a past judgment covers the current situation. Roll Wits + Lore (DV 3). On a success, the Rite's DV is reduced by 1 (minimum 1). On a miss, the precedent is contested—GM gains 1 SB.

The law remembers. The question is whether it remembers correctly.

\subsubsection{Gavel's Authority (Mykkiel) — 6 XP}

\textbf{Requirement:} Mykkiel, Body 3+

Your Codex (gavel or shield) grants you authority. Once per scene, when you strike your gavel (or strike your shield with your weapon), hostile NPCs within Near range must test Resolve (DV 4) or hesitate for one exchange.

The sound of judgment is not a suggestion.

\subsubsection{Truth's Gaze (The Witness) — 6 XP}

\textbf{Requirement:} The Witness, Presence 3+

Once per session as an action, fix your gaze on a target within Near range. For the next minute, they cannot tell a direct lie to you; they may evade or refuse to answer, but any statement they make must be factually true. Mark 1 Fatigue after using this talent.

The truth does not need to be spoken. It only needs to be witnessed.

\subsubsection{Recorded Echo (The Witness) — 4 XP}

\textbf{Requirement:} The Witness

Once per session, touch a physical trace (a footprint, a torn cloth, a drop of blood) and instantly know the last significant action taken by the person who left it (e.g., "they were running," "they hid something under the floor").

The echo fades. The memory remains.

\subsection{The Threshold (Boundaries, Wards, and Passage)}

\subsubsection{Ward Sense (The Sealed Gate) — 4 XP}

\textbf{Requirement:} The Sealed Gate

You can sense the presence of any active ward, seal, or threshold within Near range. You know its general purpose (protection, exclusion, binding) and its strength (DV).

The threshold cannot hide from the one who guards it.

\subsubsection{Sentinel's Stand (The Sealed Gate) — 6 XP}

\textbf{Requirement:} The Sealed Gate, Body 3+

When you take the Defend action while holding a threshold (door, gate, bridge), you gain +1 Armor. Any enemy attempting to cross must test Body + Athletics (DV 4) or be stopped in their tracks.

You are the wall. You do not move.

\subsubsection{Shortcut (The Traveler) — 4 XP}

\textbf{Requirement:} The Traveler

Once per session when moving from one location to another (including during travel montages), declare a hidden shortcut that reduces the remaining travel clock by 2 segments. The GM may describe a minor cost or complication (e.g., "the path is steep—mark 1 Fatigue").

The road remembers those who ask politely.

\subsubsection{Road Memory (The Traveler) — 6 XP}

\textbf{Requirement:} The Traveler

You never forget a road you have walked. You can draw a perfect map of any route you have traveled, including shortcuts, hazards, and waystations.

The map is in your feet. The journey is in your bones.

\subsubsection{Comforting Presence (The Pale Shepherd) — 4 XP}

\textbf{Requirement:} The Pale Shepherd

Once per scene, when you touch a creature suffering from Fear, Guilt, or Grief, you may remove that Condition. The creature clears 1 Fatigue if the Condition was particularly heavy (GM discretion).

The Shepherd does not heal. The Shepherd accompanies.

\subsubsection{Final Witness (The Pale Shepherd) — 6 XP}

\textbf{Requirement:} The Pale Shepherd, Spirit 3+

Once per session, when a sentient creature dies within Near range, you may ask the GM one question about that creature's final thoughts or unfulfilled wishes. The GM must answer truthfully.

The dead speak once. Listen carefully.

\subsection{The Wild (Nature, Seasons, and the Hunt)}

\subsubsection{Seasonal Sense (Grimmir) — 4 XP}

\textbf{Requirement:} Grimmir

You can predict the weather for the next hour with perfect accuracy. You know when the season will turn within a day.

The wind tells you. The soil confirms.

\subsubsection{Wild Shape (Grimmir) — 6 XP}

\textbf{Requirement:} Grimmir, Body 3+

Once per session, you may assume the form of a small beast (wolf, hawk, badger) for up to one scene. While transformed, you gain +2 dice to Stealth and Perception in natural terrain, but cannot speak or use tools. Mark 1 Fatigue after using this talent.

The skin remembers what the mind forgets.

\subsubsection{Harvest Resonance (The Carrion-King) — 4 XP}

\textbf{Requirement:} The Carrion-King

Once per scene after an enemy is defeated or a significant ending occurs, you may draw a sliver of that ending's energy. Gain +1 die on your next action, or clear 1 Fatigue.

The ending is not waste. It is compost.

\subsubsection{Rot Sense (The Carrion-King) — 6 XP}

\textbf{Requirement:} The Carrion-King

You can sense decay, rot, or corruption within Near range. You know the general direction and approximate strength (minor, moderate, severe).

The smell is not in your nose. It is in your memory.

\subsubsection{Beast Form (The Savage Heart) — 6 XP}

\textbf{Requirement:} The Savage Heart, Body 3+

Once per session, you may partially transform into a predator for one scene. Choose one: gain +2 dice to Stealth (leopard/tiger), +2 dice to Intimidation (bear/lion), or +1 band to movement (wolf/hound).

The transformation is not complete. The hunger is.

\subsubsection{Alpha's Roar (The Savage Heart) — 8 XP}

\textbf{Requirement:} The Savage Heart, Presence 3+

Once per session, you may make a terrifying sound—a roar, a howl, a scream. All enemies within Near range must test Resolve (DV 4) or become Shaken (-1 die to all actions) for one exchange.

The pack hears. The prey flees.

\subsection{The Deep (Pressure, Storms, and the Abyss)}

\subsubsection{Crushing Silence (Khemesh) — 6 XP}

\textbf{Requirement:} Khemesh

Once per session as a reaction when a creature within Near range speaks or casts a spell, impose a bubble of oppressive silence around them for one exchange. They cannot speak, and any spell with a verbal component fails. Mark 1 Fatigue after using this talent.

The deep does not shout. The deep presses.

\subsubsection{Deep Sight (Khemesh) — 4 XP}

\textbf{Requirement:} Khemesh

You can see in darkness as if it were dim light. In absolute darkness, you are not blind—you perceive pressure and movement instead of light.

The dark is not empty. It is full of weight.

\subsubsection{Storm Caller (Xhak'Thul) — 6 XP}

\textbf{Requirement:} Xhak'Thul

Once per scene, as a free action, you may cause a peal of thunder to roll across the zone. All enemies within Near must test Resolve (DV 3) or lose their next minor action.

The storm speaks. You are its echo.

\subsubsection{Lightning Reflexes (Xhak'Thul) — 4 XP}

\textbf{Requirement:} Xhak'Thul, Body 3+

You gain +1 die to defense rolls against ranged attacks (arrows, bolts, thrown weapons).

You see the path of the projectile before it flies.

\subsubsection{Tide-Step (Raéyn) — 6 XP}

\textbf{Requirement:} Raéyn

Once per session as a free action, you may treat any water-based difficult terrain as stable ground for one exchange. You may move across the surface of water as if it were solid for up to Near distance. Mark 1 Fatigue after using this talent.

The tide lifts. The tide demands.

\subsubsection{Storm Glass (Raéyn) — 4 XP}

\textbf{Requirement:} Raéyn

You can predict the weather for the next hour with perfect accuracy. You know when a storm is coming, how long it will last, and roughly how severe it will be.

The glass does not lie. The sea does not forgive.

\subsection{The Shadow (Secrecy, Silence, and the Unseen)}

\subsubsection{Shadow Meld (Ikasha) — 6 XP}

\textbf{Requirement:} Ikasha

Once per session as an action, step into any shadow larger than yourself and emerge from another shadow within Near range. You cannot be targeted while moving between shadows. Mark 1 Fatigue after using this talent.

The shadow is not absence. It is passage.

\subsubsection{Unseen Strike (Ikasha) — 4 XP}

\textbf{Requirement:} Ikasha

When you attack from a hidden position, your target suffers -1 die to defend. They do not see the threat until it is too late.

The first strike is the only strike.

\subsubsection{Burden Steady (The Silent Choir) — 4 XP}

\textbf{Requirement:} The Silent Choir

You may carry one additional psychological Condition without stacking penalties. The weight is not less—you have simply learned to bear it.

Silence is not emptiness. It is capacity.

\subsubsection{Calm Presence (The Silent Choir) — 6 XP}

\textbf{Requirement:} The Silent Choir, Spirit 3+

Once per scene, when you touch a creature suffering from Fear, Guilt, or Despair, you may remove that Condition. The creature clears 1 Fatigue if the Condition was particularly heavy.

The hand steadies. The heart follows.

\subsubsection{Forgotten Face (Aveh) — 4 XP}

\textbf{Requirement:} Aveh

Witnesses suffer -2 dice to recall your description. You are not invisible—merely unmemorable.

The storm erases. The wind forgets.

\subsubsection{Escape Clause (Aveh) — 8 XP}

\textbf{Requirement:} Aveh, Tier II+

Once per session, when you are cornered or capture is inevitable, you may declare an escape route that was not there before. The GM must accept it, but may impose a cost—you left something behind, or someone else paid the price for your freedom.

The road opens. The price is never small.

\subsection{The Web (Fate, Binding, and Transformation)}

\subsubsection{Tug the Strand (Inaea) — 4 XP}

\textbf{Requirement:} Inaea

Once per session, name a relationship between two NPCs (friend, rival, debtor, lover). The GM must reveal one secret about that relationship—a hidden grievance, a secret loyalty, an unpaid debt.

The thread is there. You only need to pull.

\subsubsection{Binding Presence (Inaea) — 6 XP}

\textbf{Requirement:} Inaea, Presence 3+

Once per session, you may compel a promise from a target unless they pass a Resolve test (DV 4). The promise is not magical—but breaking it will be noticed.

The spider does not chase. The spider waits.

\subsubsection{Shed Skin (Isoka) — 4 XP}

\textbf{Requirement:} Isoka

Once per session when you are Bound, Grappled, or suffering from a Condition that restricts movement, you may spend an action and mark 1 Fatigue to shed that restriction. Describe how you slip free like a snake shedding its skin.

The old self is a husk. The new self is already moving.

\subsubsection{Venomous Strike (Isoka) — 6 XP}

\textbf{Requirement:} Isoka

When you successfully strike a target with a melee attack, you may choose to deliver a dose of venom. The target suffers +1 Harm and must test Resolve (DV 4) or gain the Poisoned Condition.

The fang is hollow. The venom is patient.

\subsubsection{Borrowed Scars (Malachai) — 6 XP}

\textbf{Requirement:} Malachai

Once per session, transfer a Condition or 1 Fatigue from an ally to yourself. In exchange, gain +1 die on your next roll against the source of that affliction. Mark 1 Obligation after using this talent.

The coin is warm. The price is never small.

\subsubsection{Thorn's Memory (Malachai) — 4 XP}

\textbf{Requirement:} Malachai

You remember every curse you have inflicted. You can recall the face, the name, and the exact terms of any bargain you have struck.

The ledger is in your scars. The interest compounds daily.

\subsection{The Forge (Creation, Iteration, and Refinement)}

\subsubsection{Rapid Animus (Clockwork Monad) — 4 XP}

\textbf{Requirement:} Clockwork Monad

You may create a temporary infusion as a reaction (e.g., when someone draws a weapon). The infusion lasts only for the current exchange, but costs +0 Obligation (free). You may use this once per scene.

The iteration is immediate. The improvement is fleeting.

\subsubsection{Self-Optimising (Clockwork Monad) — 6 XP}

\textbf{Requirement:} Clockwork Monad, Body 3+

Once per session, you may automatically improve one of your abilities or possessions by one step for the remainder of the scene. Describe the improvement—a sharper edge, a lighter weight, a more efficient motion.

The machine learns. The machine improves. The machine never rests.

\subsection{The Grave (Ancestors, Debt, and Memory)}

\subsubsection{The Silent Ledger (Varnek Karn) — 6 XP}

\textbf{Requirement:} Varnek Karn

Once per session, touch a physical trace of a death (a body, a grave, an object that witnessed a passing) and instantly know one unresolved obligation connected to that death—a promise unfulfilled, a secret unspoken, a debt unpaid. Mark 1 Fatigue after using this talent.

The dead do not forget. They only wait.

\subsubsection{Cord of Flesh (The High Mother) — 8 XP}

\textbf{Requirement:} The High Mother, Body 3+

Your body stores memories and debts. When you convert Harm → Fatigue, you may store up to Body rating points of Fatigue as Chained Knots, appearing uninjured. You have until the sunset after the next to perform a sacrifice or take 2 Harm.

The flesh remembers. The debt accrues.

\subsubsection{The Red Thread (The High Mother) — 4 XP}

\textbf{Requirement:} The High Mother

Once per scene, you may tie a small knot in a cord, thread, or cloth within reach. For the rest of the scene, you and allies within Near range gain +1 die to a single type of roll (choose when you tie the knot). The knot must remain tied; if undone, the benefit ends.

The thread holds. The promise binds.

\subsection{The Fracture (Knowledge, Paradox, and the Unknowable)}

\subsubsection{Glimpse Beyond (The Ninth) — 4 XP}

\textbf{Requirement:} The Ninth

Once per session, look at a complex situation (a crime scene, a puzzle lock, a battle formation) and instantly intuit the single most important hidden detail. The GM gives you one piece of information that would otherwise require a successful Investigation roll. Mark 1 Fatigue after using this talent.

The pattern is there. You only need to see it.

\subsubsection{Pattern Recognition (The Ninth) — 6 XP}

\textbf{Requirement:} The Ninth

Once per scene, see connections others miss within tangled accounts and signs. You gain +1 die to any roll involving pattern recognition, system analysis, or prediction.

The threads are not random. They are merely unread.

\subsubsection{Quiet Step (Lunara) — 6 XP}

\textbf{Requirement:} Lunara

+1 die to stealth or subtle speech acts under moonlight. The silver light hides as much as it reveals.

The moon does not judge. The moon witnesses.

\subsubsection{Seam-Walker (Lunara) — 8 XP}

\textbf{Requirement:} Lunara, Wits 3+

Once per scene, you may cross a minor threshold unseen—a door, a window, a gate, a bridge. Those watching must test Resolve (DV 4) to notice your passage. Mark 1 Fatigue after using this talent.

The seam is there. You simply step through.

\subsection{The Lethai Ghe'hai (Elite Operatives)}

\subsubsection{Ghe'hai's Escape (Lethai Syncretic) — 6 XP}

\textbf{Requirement:} Lethai syncretic tradition

Once per session, when you are cornered or capture is inevitable, you may vanish, reappearing in a safe location of the GM's choosing. You leave behind a token—a silk mask, a red thread, a single word—that can be used to track you. The GM gains 2 SB.

The mask is abandoned. The mission continues.

\subsubsection{The Unseen Blade (Lethai Syncretic) — 4 XP}

\textbf{Requirement:} Lethai syncretic tradition

When you attack from a hidden position or from a crowd, your target suffers -1 die to defend. They do not see the threat until it is too late. If you are disguised, the target cannot describe you afterward.

The blade does not exist until it strikes.

\subsubsection{Silk Memory (Lethai Syncretic) — 8 XP}

\textbf{Requirement:} Lethai syncretic tradition, Witnessed Scar

Once per session, you may touch a silk cord or mask and recall a perfect sensory memory of any scene you have witnessed while wearing it—sounds, smells, temperatures, the exact words spoken. The memory may be used as testimony (in a court that respects Lethai witness) or as a clue.

The silk remembers. The mask testifies.

The thread held. The water carried. That is not a Rite. That is grace.
