\section{Rites of the Web (Inaea, Isoka, Malachai)}

\subsection{Inaea — Angel of the Spider}

\subsubsection{Low Rite: Rite of the Hearth-Thread Knot (4 XP)}

\textbf{Tags:} [BOND], [PROTECT], [HEAL], [WITNESS]

\textbf{Threshold:} A thread tied around a wrist or a shared cup.

Bind yourself and one willing target with a sympathetic thread. Both gain +1 die to Aid each other this scene. If either takes Harm, the other may reduce it by 1 (to a minimum of 0) by suffering that same Harm instead. This transfer is voluntary for an ally; for an unwilling target, you must succeed on a Resolve test (DV 3) to force the link.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Strand of Inevitability (8 XP)}

\textbf{Tags:} [BOND], [FATE], [MANIPULATION], [WITNESS]

\textbf{Threshold:} Three coloured threads (white, black, red) braided together.

Link two actors with a visible or invisible strand. For the rest of the scene, whenever one acts, the other is dragged into consequence. You may trigger this linkage once to either force 1 SB (Hearts) on the linked target after they perform a significant action, or gain +2 dice on a Setup, Predict, or Manipulate roll that exploits the link. The link is one-way or two-way as you choose.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Spinner's Debt (9 XP)}

\textbf{Tags:} [VOW], [DEBT], [BOND], [COMMAND], [WITNESS]

\textbf{Threshold:} A silver coin and a strand of your own hair, knotted together.

Offer a willing creature a boon: they may immediately clear 1 Fatigue or gain +1 die on their next roll. In return, they accept a Promise Clock [4] named "Debt to the Web." You may advance this clock by 1 segment at any time as a free action (once per scene). When the clock fills, you may demand a single service from them—not impossible, but significant (e.g., "speak for me at court," "reveal a secret," "take a blow meant for me"). Refusal causes them to suffer Harm 1 (Stress) and the clock resets.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Merciful Severing (13 XP)}

\textbf{Tags:} [HEAL], [DISPEL], [MERCY], [DEBT], [WITNESS]

\textbf{Threshold:} A white thread burned to ash, spoken with a name.

Sever one harmful tie—a curse, a toxic obsession, a coercive bond, or a supernatural debt. Remove one Condition or 1 SB tied to it. This is an act of genuine mercy; the target may be unwilling, but the thread can still be cut (Resolve test DV 4 to resist).

\textbf{Obligation:} 3

\subsection{Isoka — Angel of Serpents}

\subsubsection{Low Rite: Rite of the Loosening Skin (4 XP)}

\textbf{Tags:} [HEAL], [ESCAPE], [CLEANSE]

\textbf{Threshold:} A discarded snakeskin or a strand of your own hair.

Shrug off a temporary burden. Remove one Condition (Fear, Poison, Fatigue, or a minor social stain) that you have acquired this scene. Alternatively, reroll one failed roll to escape restraint or pursuit.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shedded Self (8 XP)}

\textbf{Tags:} [TRANSFORM], [ERASURE], [ESCAPE], [DEBT]

\textbf{Threshold:} A complete change of clothing and a token from your former role (a badge, a lock of hair, a written name).

You shed one persistent identity—a title, a bond, a debt, or a reputation. For the remainder of the scene, you are treated as a stranger to those who knew you in that role. You gain +2 dice on any roll to avoid detection, escape pursuit, or convince others that you are not yourself. Create a 6-segment Shed Identity clock; when it fills, the old identity returns as a separate NPC (a rival, a witness, or a ghost).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of Venom's Paradox (9 XP)}

\textbf{Tags:} [POISON], [HEAL], [CURSE], [CLEANSE], [WITNESS]

\textbf{Threshold:} A drop of venom from a serpent that has shed its skin within the last day.

Inflict or cure through poison's paradox. Choose one: cursed venom (touch a creature and inject venom; they suffer +1 Harm and must test Resolve (DV 4) or gain the Poisoned Condition) or mercy's antidote (touch a poisoned or diseased creature; the venom neutralises the malady; remove one Poisoned or Diseased Condition, and the target clears 1 Fatigue).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of Unremembered Metamorphosis (13 XP)}

\textbf{Tags:} [TRANSFORM], [ERASURE], [FATE], [BURDEN]

\textbf{Threshold:} A full identity kit, a serpent's shed skin, and a personal item from the life you are leaving (or records of your old life to burn).

Choose one: complete metamorphosis (after an hour of ritual, you emerge with a new face, voice, and documented identity; for one week, immune to recognition by anyone who knew your previous self, and gain +2 dice on all Deception and Stealth rolls related to that new identity) or cast-off history (after a full night's ritual burning all mundane records of your old life, your previous identity becomes unreliable; common records and casual memories become vague; you gain +2 dice on rolls relying on your new identity).

\textbf{Obligation:} 3

\subsection{Malachai — The Cruel Messenger}

\subsubsection{Low Rite: Rite of the Honeyed Binding (5 XP)}

\textbf{Tags:} [CURSE], [BOOST], [DEBT], [FEAR], [WITNESS]

\textbf{Threshold:} Blood mixed with sweet wine, or a broken chain link.

Choose one: honeyed curse (target gains +2 dice to their next roll—counts as Triumph; they take 1 Corruption, unresistable) or binding curse (target gains +1 die to physical actions but suffers Fatigue 1 unless they perform a dominance act—beat a weaker creature, drink blood, break something precious).

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the False Dawn (8 XP)}

\textbf{Tags:} [CURSE], [DEBT], [FATE], [WITNESS]

\textbf{Threshold:} Meteorite iron shard blessed by false light.

Bestow a gift with a hidden curse. Target gains a significant advantage (e.g., +1 die to all rolls for one scene) but accrues a 4-segment Corruption clock. Each time they use the gift, tick the clock. When full, they suffer a related curse (GM chooses: festering wounds, beastly urges, or the attention of a Collector).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Cruel Transformation (9 XP)}

\textbf{Tags:} [CURSE], [TRANSFORM], [DEBT], [FEAR], [WITNESS]

\textbf{Threshold:} Holy water mixed with graveyard dust.

Infect a target with a supernatural condition (vampiric hunger, bestial rage, shadow-bind). Lasts one scene. Start a 6-segment Transformation clock. Each time the target resists feeding or transformation, tick the clock. When full, the condition becomes permanent.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Chained Angel's Tribute (14 XP)}

\textbf{Tags:} [CURSE], [AREA], [SACRIFICE], [FATE], [WITNESS]

\textbf{Threshold:} A chapel bell cracked by divine lightning, or a silver coin that cannot be spent.

Corrupt a sacred space or consecrate a site to Malachai. Within the zone: supernatural effects (curses, dark magic) gain +1 Effect; blessings and healing spells generate +1 SB on a failure; any creature that accepts a gift from Malachai within the zone immediately ticks a Corruption clock twice. Start a 10-segment Divine Perversion clock. When it fills, the site becomes a permanent beacon for the Worn—the ghouls of Zahk'Kozahn.

\textbf{Obligation:} 3

\subsection{Rite Synergies}

Rites from the Web family complement each other in sequence:

\begin{itemize}
\item \textbf{Inaea's Strand of Inevitability + Isoka's Shedded Self:} Bind two actors, then shed your identity. The link remains—but the target cannot find you to demand reciprocity.
\item \textbf{Malachai's Honeyed Binding + Isoka's Venom's Paradox:} Curse an enemy with borrowed strength, then offer them the antidote—for a price. The poison and the cure come from the same hand.
\item \textbf{Inaea's Merciful Severing + Malachai's Cruel Transformation:} Cut a curse you inflicted—or let it bloom. The thread is yours to tighten or release.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
