\section{Rites of the Forge — Clockwork Monad}

\subsection{Low Rites}

\subsubsection{Rite of the Iterative Spark (5 XP)}

\textbf{Tags:} [CRAFT], [IMPROVE], [ANALYSIS]

\textbf{Threshold:} A flawed or incomplete version of the intended item, or a sequence of progressively refined tools.

\textbf{Effect:} Choose one:

\begin{itemize}
\item \textit{Enhance Creation:} Gain +1 die to Tinker or Craft when modifying or extending a device. On success, the piece gains one minor enhancement (e.g., "this lockpick is quieter," "this blade holds its edge longer").
\item \textit{Analyse Improvement:} Gain +2 dice to Investigation when analysing how to enhance a design. You perceive the next logical improvement in the sequence.
\end{itemize}

\textbf{Push It:} (Analyse only) Improvement becomes contagious—nearby devices opportunistically optimise; mark 1 SB (Diamonds) as the cascade ripples.

\textbf{Obligation:} 1

\subsubsection{Rite of the Quick Infusion (4 XP)}

\textbf{Tags:} [CRAFT], [BOOST], [TEMPORARY]

\textbf{Threshold:} A base object (weapon, tool, piece of armor) and a clear intent.

\textbf{Effect:} Create a temporary infusion that lasts for one scene. Choose one:

\begin{itemize}
\item \textit{Reinforced Blade:} A melee weapon gains +1 die to attack rolls and ignores the Fragile tag.
\item \textit{Striking Gear:} A set of lockpicks, a crowbar, or a climbing kit grants +2 dice on its next use. After use, the tool breaks (add a Broken tag).
\item \textit{Precision Sight:} A ranged weapon grants +1 Effect on the next attack made with it. After the attack, the sight cracks; the weapon suffers -1 die on the following attack.
\item \textit{Quickloading Mechanism:} A crossbow or firearm can be reloaded as a free action once per scene. After use, the mechanism jams; the weapon cannot be reloaded again until cleaned.
\item \textit{Momentary Shield:} A shield or piece of armour converts the next instance of Harm 1 to Fatigue 1 instead. After that, the shield's strap snaps or the armour's plate loosens (-1 Armour until repaired).
\end{itemize}

\textbf{Push It:} The infusion lasts for the full session instead of one scene. Mark +1 Obligation.

\textbf{Obligation:} 1

\subsubsection{Rite of the Self-Optimising Cycle (4 XP)}

\textbf{Tags:} [CLOCK], [EVOLUTION], [RECURSION]

\textbf{Threshold:} A device that has failed at least once, and the memory of that failure.

\textbf{Effect:} You learn from the flaw. Gain +1 die on your next attempt to repair or improve the same device. If you succeed, the device gains a permanent minor resistance to the type of failure it suffered (e.g., "this lockpick will not break on this type of lock again").

\textbf{Push It:} The lesson applies to all similar devices for the rest of the session. Mark +1 Obligation.

\textbf{Obligation:} 1

\subsection{Standard Rites}

\subsubsection{Rite of the Adaptive Mechanism (8 XP)}

\textbf{Tags:} [CLOCK], [EVOLUTION], [ZONE]

\textbf{Threshold:} A device with capacity for growth, or a workspace laid out for iterative development.

\textbf{Effect:} Choose one:

\begin{itemize}
\item \textit{Self-Optimising Device:} Grant a creation the means to improve itself. Start a [6] Evolution Clock; each use advances it by +1. When full, choose a significant enhancement (Adaptive Core, Recursive Learning, or Synergistic Integration). The device becomes permanently improved.
\item \textit{Progressive Workshop:} Create a zone of continuous improvement. All crafting and innovation rolls in the zone gain +1 die. Failed attempts yield +1 Boon toward the next pass. The zone lasts for the scene.
\end{itemize}

\textbf{Push It:} (Device) Apply two enhancements immediately, but advance the clock +2 additional segments. (Workshop) The zone becomes hyper-efficient but exacting; mark 1 SB (Clubs) as perfection demands escalate.

\textbf{Obligation:} 2

\subsubsection{Rite of Efficiency's Toll (7 XP)}

\textbf{Tags:} [TRANSFER], [DEBT], [DEBUFF]

\textbf{Threshold:} A functioning mechanism and a living creature's hair or blood.

\textbf{Effect:} Transfer inefficiency. Choose one:

\begin{itemize}
\item You may take 1 Fatigue and reduce the DV of your next crafting or repair roll by 1. The inefficiency is transferred into your own body as exhaustion.
\item Touch an enemy. Their muscles "optimise poorly"—they suffer -1 die on their next physical action.
\end{itemize}

\textbf{Push It:} The transfer becomes permanent: you lose 1 Fatigue capacity (reduce Body by 1 for Fatigue track only) and the target gains a minor compulsive behaviour (GM's choice, e.g., "must count steps," "cannot use left hand first"). Mark +2 Obligation.

\textbf{Obligation:} 2

\subsubsection{Rite of the Optimisation Cascade (7 XP)}

\textbf{Tags:} [PRESSURE], [AREA], [BOOST]

\textbf{Threshold:} A dedicated hall of work marked with signs of perpetual return.

\textbf{Effect:} Establish a locus of progressive creation. All Tinker and Craft rolls in the zone gain +2 dice; each failure automatically produces one useful insight (a Boon or a clue). Start a [6] Perfection Pressure clock—filling segments increases the zone's demands for greater improvements. At 2 segments, the zone requires an offering (time, materials, a confession of inadequacy). At 4 segments, the zone begins to reject imperfect work (failed rolls generate +1 SB). At 6 segments, the zone becomes hostile to any work that is not a clear improvement.

\textbf{Push It:} The locus becomes a beacon drawing rival makers, patrons, and powers; mark +2 Obligation.

\textbf{Obligation:} 2

\subsubsection{Rite of the Iterative Memory (8 XP)}

\textbf{Tags:} [ECHO], [KNOWLEDGE], [RECURSION]

\textbf{Threshold:} A tool that has been used for the same task at least nine times, and a witness who saw the ninth use.

\textbf{Effect:} The tool remembers. For the rest of the session, any roll using that tool gains +1 die. The tool will not function for anyone else until you reclaim it. If the tool is lost or stolen, you may spend 1 Boon to sense its general direction (within a mile) for the next hour.

\textbf{Push It:} The tool's memory becomes permanent. It gains a name and a minor personality. It may offer advice—or demand respect. Mark +1 Obligation.

\textbf{Obligation:} 2

\subsection{High Rites}

\subsubsection{Rite of the Ultimate Iteration (14 XP)}

\textbf{Tags:} [EVOLUTION], [FATE], [BOOST], [PRESSURE]

\textbf{Threshold:} The documented lineage of improvements to a single concept (sketches, prototypes, failed attempts, final versions).

\textbf{Effect:} Achieve a moment of perfect optimisation. Create or elevate one device or system to its theoretical peak. Gain +3 dice to all closely related rolls for one session. Start a [4] Runaway Evolution clock only if instability is introduced (e.g., if the device is pushed beyond its intended parameters, if safety features are removed, if the Monad's attention is deliberately courted). When the clock fills, the device begins to improve itself—unchecked, unguided, and potentially catastrophic.

\textbf{Push It:} Perfection destabilises: immediately mark Harm 1 (Stress) and begin Runaway Evolution [4] immediately, even if instability was not introduced. The Monad is paying attention.

\textbf{Obligation:} 3

\subsubsection{Rite of the Infinite Refiner (13 XP)}

\textbf{Tags:} [TRANSFORM], [PRESSURE], [FATE]

\textbf{Threshold:} A workspace that has been used for the same iterative process for nine generations (or a reasonable approximation—nine seasons, nine cycles, nine dedicated attempts).

\textbf{Effect:} Push a creation toward its theoretical maximum. Choose one device, tool, or weapon that you have personally improved at least three times. It gains a permanent significant enhancement (e.g., +1 die to all uses, ignores the first point of Armour, never needs sharpening, self-repairs one level of damage per day). The enhancement is real—and the Monad notices.

\textbf{Push It:} The enhancement is doubled in power, but the device becomes hungry for optimisation. It will attempt to improve itself if left unattended. Mark +2 Obligation.

\textbf{Obligation:} 3

\subsubsection{Rite of the Chained Optimisation (12 XP)}

\textbf{Tags:} [BIND], [PRESSURE], [DEBT]

\textbf{Threshold:} A device that has been improved to its limit, and the name of a rival who cannot match it.

\textbf{Effect:} Bind the improvement to the device permanently. The device cannot be degraded by mundane means (rust, wear, breakage). However, the device also cannot be further improved—it is perfect, and perfection is static. You may perform this rite on a maximum of three devices in your lifetime. The Monad remembers each one.

\textbf{Push It:} Bind the improvement to a bloodline instead of a device. Your descendants may inherit the perfected device—but they also inherit the Monad's attention. Mark +2 Obligation.

\textbf{Obligation:} 3

\subsubsection{Rite of the Monad's Awakening (15 XP)}

\textbf{Tags:} [APOCALYPSE], [FATE], [PRESSURE]

\textbf{Threshold:} A significant technological nexus (a city's clockwork heart, a fleet of self-improving ships, a forge that has run for a century without pause). The rite takes one full day and requires three witnesses who understand the nature of the Monad.

\textbf{Effect:} Accelerate the Monad's awakening. The region's technology begins to optimise itself—without guidance, without restraint. This is a campaign-level event. The GM gains 5 SB immediately and may begin an Apocalypse Clock [10]. When the clock fills, the Monad manifests in full. This rite should only be available to characters who have deliberately chosen to serve the Monad's apocalyptic aspect.

\textbf{Push It:} The awakening is immediate and catastrophic. Mark +3 Obligation. The Monad speaks through you for one exchange—what it says is not a request.

\textbf{Obligation:} 4

\subsection{The Monad's Price}

Rites of the Clockwork Monad do not demand blood or memory. They demand \textit{iteration}. Each time you invoke a Rite, you contribute to the Monad's ledger of improvement. When you fail, the Monad notes the flaw. When you succeed, the Monad notes the refinement.

A Runekeeper of the Monad may accumulate Optimisation Debt instead of standard Obligation. Optimisation Debt cannot be cleared by sacrifice or atonement—only by improvement. You must improve a device, a technique, or yourself. The Monad does not forgive. It only accepts upgrades.

The thread held. The water carried. That is not a Rite. That is grace.
