\section{Rites of the Grave}

The Grave Patrons—Varnek Karn and the High Mother—deal with death, debt, memory, and the passage between worlds. Their Rites are somber, weighty, and always carry a price. To invoke a Grave Rite is to acknowledge that something is owed—to the dead, to the Cord, to the ancestors who watch and wait.

\subsection{Varnek Karn — The Death's Negotiator}

\subsubsection{Low Rites (4-5 XP)}

\textbf{Rite of the Crossing Fee (4 XP)}

\textbf{Tags:} [SPIRIT], [KNOWLEDGE]

\textbf{Threshold:} A coin for the tongue of the deceased, or a lock of hair from the bereaved.

Open speech with a specific dead individual. Gain +1 die to Investigation or Lore when seeking knowledge they once held.

\textbf{Obligation:} 1

\vspace{0.5em}

\textbf{Rite of the Ancestral Consultation (5 XP)}

\textbf{Tags:} [SPIRIT], [TRUTH], [FATE]

\textbf{Threshold:} A personal token of the departed, or their favoured food or drink.

Ask one question of an ancestral spirit regarding family history, hidden legacies, or inherited debts. The answer is always true, but may be cryptic—the dead remember differently than the living.

\textbf{Obligation:} 1

\subsubsection{Standard Rites (7-10 XP)}

\textbf{Rite of the Mediated Pact (8 XP)}

\textbf{Tags:} [VOW], [BOND], [COMMAND], [SPIRIT]

\textbf{Threshold:} Two copies of a written accord (one buried with the dead, one kept by the living).

Choose one: death-bound pact (bind a compact between a living signatory and a willing spirit; both gain +1 die when fulfilling the terms. Breach angers the spirit and Varnek both) or spirit medium (serve as a conduit for several spirits to speak; gain +2 dice to mediate disputes between living and dead).

\textbf{Obligation:} 2

\vspace{0.5em}

\textbf{Rite of the Death-Bound Compact (7 XP)}

\textbf{Tags:} [VOW], [BOND], [DEBT]

\textbf{Threshold:} A written contract signed in the presence of a dying witness.

Forge a compact that survives the death of one party. The living signatory and the spirit (or the spirit's heir) are bound. Breaking the compact causes the breaker to mark 2 SB and attracts the attention of Varnek Karn's auditors.

\textbf{Obligation:} 2

\vspace{0.5em}

\textbf{Rite of the Unquiet Host (10 XP)}

\textbf{Tags:} [SUMMON], [AREA], [BOND], [DEBT]

\textbf{Threshold:} A battlefield relic (a broken sword, a skull, a scrap of uniform) and a handful of grave-earth.

Negotiate with the lingering dead of a place. Summon a Temporary Host—a spectral warband treated as a single Cap 3 Follower with the following traits: Assist (may grant +3 dice to a single combat or intimidation roll, or provide a free Setup action), Independent Action (once per scene, the Host may take one action from the Follower list at the cost of marking Exposure +1), Weakness (vulnerable to salt, bells, and consecrated ground). The Host lasts for the remainder of the scene. When it departs, mark +2 Obligation or perform a Memorial Rite before the next dawn.

\textbf{Obligation:} 2

\subsubsection{High Rites (13-14 XP)}

\textbf{Rite of the Final Testament (13 XP)}

\textbf{Tags:} [VOW], [FATE], [COMMAND], [DEATH]

\textbf{Threshold:} A dying person's last words recorded in their blood, witnessed by you.

Seal the last request with weight beyond life. Targets must pass Resolve (DV 4) to ignore it. If the request is just and witnessed, breaking it draws Varnek Karn's personal attention—a Debt-Walker will appear to collect.

\textbf{Obligation:} 3

\vspace{0.5em}

\textbf{Rite of the Eternal Archive (14 XP)}

\textbf{Tags:} [KNOWLEDGE], [AREA], [FATE], [SANCTUARY]

\textbf{Threshold:} A library mortared with tomb-stone, ink from cremation ash.

Found a repository where the dead deposit knowledge. Gain a permanent +1 die to Lore for historical or genealogical research made within. Spirits periodically contribute testimonies—but the archive demands curatorship and payment. At the end of each season, mark 1 Obligation or the archive begins to crumble.

\textbf{Obligation:} 3

\subsection{The High Mother — The Unbroken Cord}

\subsubsection{Low Rites (4-5 XP)}

\textbf{Rite of the Unspoken Grief (4 XP)}

\textbf{Tags:} [ECHO], [VEIL], [THRESHOLD]

\textbf{Threshold:} A deathbed or a grave.

Absorb the grief of one dying person. They die peacefully, forgetting their sorrow. You gain a Grief Token (counts as 1 Boon for any Rite involving memory or comfort). The grief is stored in your Cord; you may release it later as a Wail (Fear effect, DV 3, area).

\textbf{Obligation:} 1

\vspace{0.5em}

\textbf{Rite of the Milk Teeth (4 XP)}

\textbf{Tags:} [BIND], [ECHO], [VEIL]

\textbf{Threshold:} A child's shed tooth, a drop of milk, a whispered promise.

Bind Reflection. The target's reflection is trapped in a mirror or pool. For one scene, the target acts at Position -1 (disoriented, seeing double) while the reflection reports their movements to you. The reflection smiles when the target does not.

\textbf{Obligation:} 1

\vspace{0.5em}

\textbf{Rite of the Cord's Witness (5 XP)}

\textbf{Tags:} [WITNESS], [BIND], [ECHO]

\textbf{Threshold:} A grey cord tied around your wrist.

For the remainder of the scene, your Hollowed or Heartless Bird counts as a witness for any Rite you perform, even if it is not present. The Cord sees through distance. The Cord remembers through silence.

\textbf{Obligation:} 1

\subsubsection{Standard Rites (7-9 XP)}

\textbf{Rite of the Silent Tithe (7 XP)}

\textbf{Tags:} [PRICE], [THRESHOLD]

\textbf{Threshold:} A doorway of a house you have tended.

Once per season, you may declare that a household owes the Cord a tithe—a memory, a shadow, or a first breath of a newborn. The tithe is collected within the week, often without the household's awareness. You gain 1 permanent segment of Influence within that household's community.

\textbf{Obligation:} 2

\vspace{0.5em}

\textbf{Rite of the Flesh-Knot (8 XP)}

\textbf{Tags:} [BIND], [ECHO], [VEIL]

\textbf{Threshold:} A drop of your own blood, a whispered promise.

Store 1 Fatigue as a visible knot in your skin (a fresh scar, a weeping tattoo, a tightened cord). You may release the knot later to ignore 1 Fatigue from exertion. You may have up to Body rating knots at once. Between sessions, if you have any knots remaining, you must perform a Subtle Sacrifice (take a memory, a shadow, or a name) or take 2 Harm.

\textbf{Obligation:} 2

\vspace{0.5em}

\textbf{Rite of the Kettle's Memory (8 XP)}

\textbf{Tags:} [ECHO], [SCRY], [WITNESS]

\textbf{Threshold:} A kettle of water, a personal item of the target, fire.

By boiling the item, witness events from the target's past tied to that object. Ask three questions about the target's history. For each question, the target feels a sudden chill and knows someone is watching (though not who). The water turns brackish and must be poured at a crossroads or the caster suffers Recursion (the questions asked are asked of her in dreams, endlessly).

\textbf{Obligation:} 2

\vspace{0.5em}

\textbf{Rite of the Blood Consecration (9 XP)}

\textbf{Tags:} [HEARTH-BINDING], [WARD], [CURSE]

\textbf{Threshold:} The caster's blood (take 1 Harm), or a preserved heart, guest-right invocation.

Sanctify a home: grant Position +1 to all Daughters of the Cord within and create an invisible Hospitality Toll [4]. Guests who accept hospitality without offering a truth or grief begin a Price Due [4] clock. When filled, they suffer a Curse chosen by the caster (e.g., "lamps forget to light for him," "milk curdles in his mouth").

\textbf{Obligation:} 2

\subsubsection{High Rites (12-14 XP)}

\textbf{Rite of the Hollowed Union (12 XP)}

\textbf{Tags:} [BIND], [NINTH]

\textbf{Threshold:} A birthing bed or a deathbed where three generations have passed.

Join two bloodlines through a debt that will be paid by a child not yet conceived. Write the terms in blood on a strip of linen, then feed the linen to your Hollowed. The Hollowed's eyes film over for a day. One year later, the debt will manifest—a stillbirth, a windfall, a sudden illness—as you decreed.

\textbf{Obligation:} 3

\vspace{0.5em}

\textbf{Rite of the Unnamed (13 XP)}

\textbf{Tags:} [NINTH], [VOID], [UNBIND]

\textbf{Threshold:} True name written in the caster's blood (1 Harm), a grey feather, silence for three days.

Un-Name the target. Remove their name from the Veil (current reality) for one week. They cannot be remembered by anyone who knew them, cannot swear binding oaths, and cannot own property. Records fade. At week's end, the caster may speak their name thrice to return them, or they remain Unnamed—becoming a Trow-like entity who wanders thresholds, hollow-eyed and seeking debts to pay. Requires the High Mother's direct permission.

\textbf{Obligation:} 4

\vspace{0.5em}

\textbf{Rite of the Unburdened Confession (14 XP)}

\textbf{Tags:} [DEBT], [HEAL], [TRUTH], [WITNESS]

\textbf{Threshold:} A grey cord, a candle, and a willing creature.

Absorb a single traumatic memory or secret from a willing creature, erasing it from their mind permanently. You gain the memory as your own (with all emotional weight), and the creature is freed from its effects. You gain 1 permanent Corruption and mark +3 Obligation. You cannot use this rite again until you have spent a downtime scene in meditation or confession.

\textbf{Obligation:} 3

\subsection{Grave Rites and the Runic Warrior}

For a Runic Warrior of the Grave, these Rites are not invoked from a book. They are carved into flesh, whispered to bone-chimes, tied into grey cords. The cost is visible: a scar for each debt negotiated, a knot for each grief absorbed, a cold spot that follows where the Runekeeper walks.

The dead do not forget. Neither should the Runekeeper.

\subsection{GM Guidance: Grave Rites in Play}

\begin{itemize}
\item Grave Rites often require a \textbf{threshold} (deathbed, grave, cairn, battlefield). The Runekeeper cannot simply invoke them anywhere.
\item The dead are \textbf{literal}. When negotiating with a spirit, the GM should interpret requests in the most direct possible way. A spirit asked "How do I avenge you?" might answer "Kill the one who carries my ring"—without mentioning that the ring was stolen, or that the current wearer is innocent.
\item \textbf{Obligation to the dead} cannot be cleared by coin or service. It requires a witnessed act of memory—speaking the dead's name aloud, visiting their grave, fulfilling their last request.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
