\section{Rites of the Fracture}

The Fracture holds two patrons: The Ninth, who offers hyperdimensional knowledge and impossible patterns, and Lunara, who guides through moonlight and hidden thresholds. Their Rites share a common theme—revelation, passage, and the cost of seeing too clearly.

\subsection{The Ninth — Hyperdimensional Rites}

\subsubsection{Low Rite: The Unfinished Equation (4 XP)}

\textbf{Tags:} [KNOWLEDGE], [PATTERN], [STRAIN]

\textbf{Threshold:} A chalkboard or slate with an incomplete equation, or a problem that has stumped the user for at least an hour.

You glimpse the missing term. For one scene, gain +1 die to all rolls involving calculation, analysis, or system understanding. The answer is there—you simply could not see it before.

\textbf{Obligation:} 1

\textbf{Backlash:} On a miss, the equation inscribes itself behind your eyes. Mark 1 Fatigue and suffer -1 die to all non-analytical rolls for the scene (the numbers will not stop).

\subsubsection{Low Rite: The Overheld Note (5 XP)}

\textbf{Tags:} [AUDIO], [THRESHOLD], [PATTERN]

\textbf{Threshold:} A note held longer than breath—hummed, sung, or played—until the air itself seems to listen.

For one scene, you may hear conversations through walls, around corners, or across a crowded room as if you stood beside the speakers. The pattern of sound reveals what should be hidden.

\textbf{Obligation:} 1

\textbf{Backlash:} On a partial, you hear everything—including the thoughts beneath the words. Mark 1 Fatigue and gain a Condition: \textit{Noise-Blind} (-1 die to all rolls requiring focus) until you spend an hour in silence.

\subsubsection{Standard Rite: The Probability Braid (8 XP)}

\textbf{Tags:} [PREDICTION], [FATE], [PATTERN]

\textbf{Threshold:} Three strands of thread or hair, braided while reciting possible outcomes.

Before a significant roll, you may braid the threads. Choose one: gain +2 dice to the roll, or treat a Miss as a Partial Success. The braid is consumed.

\textbf{Obligation:} 2

\textbf{Backlash:} On a miss, the threads tighten around your wrist. Mark 1 Fatigue and gain a Promise Clock [4] named \textit{The Unbraided Outcome}. When it fills, the GM may force a reroll of any one of your successful rolls.

\subsubsection{Standard Rite: The Recursive Glimpse (9 XP)}

\textbf{Tags:} [PRECOGNITION], [TIME], [PATTERN]

\textbf{Threshold:} A mirror facing another mirror, creating an infinite regress, or a still pool at midnight.

You see one second into the future—enough to dodge a blow, catch a falling object, or choose the correct door. For the rest of the scene, you gain +1 die to all defense rolls and may reroll one failed roll (after seeing the result, before consequences are applied).

\textbf{Obligation:} 2

\textbf{Backlash:} On a partial, the glimpse is delayed. You see the future one second after it happens. Mark 1 Fatigue and lose the benefit of this Rite for the scene.

\subsubsection{High Rite: The Fractal Proof (14 XP)}

\textbf{Tags:} [TRUTH], [SYSTEM], [PATTERN], [REVELATION]

\textbf{Threshold:} A complete proof—mathematical, logical, or legal—that is wrong, recited aloud while walking a spiral inscribed on the floor.

You prove the unprovable. Choose one: prove that a system (legal, magical, bureaucratic) contains a contradiction (it must be revised or collapses), prove that an enemy's action was inevitable (they cannot be blamed for it—but cannot claim innocence), or prove that a hidden truth must exist (the GM must reveal one fact that was previously concealed).

The proof lasts as long as the spiral remains unbroken. If anyone steps across the line, the Rite ends.

\textbf{Obligation:} 3

\textbf{Backlash:} On a miss, the proof proves itself—and you are the contradiction. Mark 2 Harm (Stress) and gain a Condition: \textit{Fractal Scar} (the GM may force you to reroll any one successful roll per session, as reality rejects your logic).

\subsection{Lunara — Lunar and Threshold Rites}

\subsubsection{Low Rite: The Veil of Silver (4 XP)}

\textbf{Tags:} [ILLUSION], [STEALTH], [THRESHOLD]

\textbf{Threshold:} A silver coin placed over each eye (or held to the eyes) while the moon is visible, even through cloud cover.

Moonlight wraps around you like a cloak. For one scene, you gain +1 die to Stealth and Deception rolls while in dim light or darkness. You are not invisible—merely forgettable.

\textbf{Obligation:} 1

\textbf{Backlash:} On a partial, the moonlight clings too tightly. You gain +1 die to Stealth but suffer -1 die to all social rolls (others find you unsettlingly pale and still).

\subsubsection{Low Rite: The Opened Way (5 XP)}

\textbf{Tags:} [MOVEMENT], [GUIDANCE], [THRESHOLD]

\textbf{Threshold:} Three lit candles at a crossroads, or three silver coins placed at a door's threshold.

A path opens where none existed. You and your allies may ignore one level of difficult terrain or bypass one mundane obstacle (a locked door, a collapsed bridge, a confusing maze) as if it were not there. The path lasts for one exchange.

\textbf{Obligation:} 1

\textbf{Backlash:} On a partial, the path opens—but so does something else. A spirit, a creature, or a memory follows you through. The GM gains 1 SB.

\subsubsection{Standard Rite: The Watcher's Reflection (8 XP)}

\textbf{Tags:} [REVELATION], [KNOWLEDGE], [TRUTH]

\textbf{Threshold:} A mirror or still water under moonlight, with a question whispered into the surface.

The reflection shows the answer to one question about a person, place, or object within the same realm. The answer is always true—but it may be symbolic, incomplete, or shown from an angle you did not expect.

\textbf{Obligation:} 2

\textbf{Backlash:} On a partial, the reflection shows a question you did not ask. The GM reveals a secret—but it is not the one you wanted. Mark 1 Fatigue.

\subsubsection{Standard Rite: The Lunar Respite (7 XP)}

\textbf{Tags:} [HEAL], [RESTORATION], [CLEANSE]

\textbf{Threshold:} A bed or resting place touched by moonlight, and a willing sleeper.

The sleeper rests as if a full night has passed in a single hour. They clear 2 Fatigue and remove one minor Condition (Fear, Shaken, Guilty). The sleeper must remain in moonlight for the entire hour; if the light is blocked, the Rite fails.

\textbf{Obligation:} 2

\textbf{Backlash:} On a miss, the sleeper dreams of falling. They clear no Fatigue and gain the Condition \textit{Exhausted} (cannot benefit from any rest for 24 hours).

\subsubsection{High Rite: The Silver Key (13 XP)}

\textbf{Tags:} [UNLOCK], [GATE], [FATE], [THRESHOLD]

\textbf{Threshold:} A silver key forged or anointed under a new moon, and a door, gate, or threshold that has been sealed for at least a year.

Open a way where none should exist—between places, states, or wards. The key turns once. The door opens. What lies beyond is not guaranteed to be safe.

\textbf{Obligation:} 3

\textbf{Backlash:} On a partial, the door opens—but something else opens with it. A second door, behind the first. The GM gains 2 SB, and the key cracks (becomes Compromised until repaired).

\subsubsection{High Rite: The Moondown Vigil (12 XP)}

\textbf{Tags:} [WARD], [PROTECT], [SANCTUARY]

\textbf{Threshold:} A circle of silver coins or moonlit stones, drawn at dusk and maintained until dawn.

Consecrate a zone against those who hate the light. Undead, lycanthropes, and creatures of shadow cannot enter the circle without testing Resolve (DV 5). Mortals who enter with ill intent suffer -1 die to all actions while within.

The circle lasts until dawn or until broken. If the moon sets before dawn (as it does), the Rite ends early—the darkness has no mercy.

\textbf{Obligation:} 3

\textbf{Backlash:} On a miss, the circle attracts the very things it was meant to ward. A creature of shadow appears at its edge. It is not hostile—yet. It is curious. It asks a question. Answer poorly, and it will enter.

\subsection{Shared Rites (Both Patrons)}

\subsubsection{Low Rite: The Glimpse (4 XP, shared)}

\textbf{Tags:} [PRECOGNITION], [PATTERN], [STRAIN]

\textbf{Threshold:} A held breath, a closed eye, a moment of perfect stillness.

You see one second into the future—enough to know whether a die will succeed or fail before you roll. Declare you are using this Rite before a roll. The GM tells you "success" or "failure." You may then choose to roll anyway (with the knowledge) or take a different action.

\textbf{Obligation:} 1

\textbf{Backlash:} On a miss, the glimpse is wrong. The GM tells you the opposite of the truth. Mark 1 Fatigue and suffer -1 die on the roll if you proceed.

\subsubsection{Standard Rite: The Unseen Pattern (7 XP, shared)}

\textbf{Tags:} [INVESTIGATION], [PATTERN], [TRUTH]

\textbf{Threshold:} A complex scene (a crime scene, a battlefield, a bureaucratic archive) and a question about its hidden structure.

You see the pattern beneath the chaos. Ask the GM one question about the scene's hidden logic (e.g., "What connects these three deaths?" "Where is the flaw in this contract?" "Who benefits from this confusion?"). The GM must answer truthfully, though the answer may be cryptic.

\textbf{Obligation:} 2

\textbf{Backlash:} On a miss, the pattern sees you. Mark 1 Fatigue and gain a Condition: \textit{Pattern-Touched} (the GM may, once per session, show you a vision of a pattern that does not exist—and you must act as if it were real until proven otherwise).

\subsubsection{High Rite: The Revealed Architecture (14 XP, shared)}

\textbf{Tags:} [REVELATION], [SYSTEM], [FATE], [TRUTH]

\textbf{Threshold:} A complete map or diagram of the system you wish to understand—drawn by your own hand, with every detail correct, under moonlight or candlelight.

You see the system as it truly is—not as it claims to be. For the rest of the session, you gain +2 dice to all rolls that exploit the system's hidden rules, contradictions, or blind spots. You may also ask the GM one question about the system's creators, purpose, or ultimate flaw; the GM must answer truthfully.

\textbf{Obligation:} 3

\textbf{Backlash:} On a partial, you see the system—and the system sees you. Mark 2 Fatigue and gain a Promise Clock [6] named \textit{System Observation}. When it fills, the system's guardians (or the system itself) come to investigate.

The thread held. The water carried. That is not a Rite. That is grace.
