\section{Rites of the Lethai Ghe'hai}

The Lethai do not have patrons in the same way as other peoples. Their magic is woven into the fabric of their society—silk, shadow, and the careful negotiation of thresholds. The Ghe'hai, the elite operatives of Lethai-ar and Lethai-thora, draw on syncretic traditions that blend Inaea's webs, Ikasha's shadows, Isoka's transformations, and the Silent Choir's discretion. Their Rites are subtle, precise, and almost never overt.

\subsection{Core Principle: The Witnessed Act}

Ghe'hai Rites require a witness—not necessarily a person. A spider's web, a still pool, a silver mirror, or a Hollowed bird can serve. The witness does not need to understand what it sees. It only needs to see.

When a Ghe'hai Rite is performed in the presence of a witness (including a sentient Thiasos or Hollowed), reduce the Obligation cost by 1 (minimum 0). The web remembers.

\subsection{Low Rites (4-5 XP)}

\subsubsection{Rite of the Silk Binding (4 XP)}

\textbf{Tags:} [BIND], [STEALTH], [THRESHOLD]

\textbf{Threshold:} A strand of spider-silk, a length of red thread, or a strand of your own hair.

Immobilize a single target with invisible threads. The target must test Body + Athletics (DV 4) to move more than a step in any direction. The binding lasts for one exchange, or until the target succeeds on the test.

\textbf{Obligation:} 1

\textbf{Backlash:} On a miss, the threads bind you instead. You are immobilized for one exchange. The GM gains 1 SB.

\subsubsection{Rite of the Forgotten Face (5 XP)}

\textbf{Tags:} [VEIL], [DECEPTION], [STEALTH]

\textbf{Threshold:} A mask, a veil, or a handful of dust thrown into the air.

Your features blur. For one scene, any witness who sees your face must test Resolve (DV 3) or forget your features within the hour. They will remember meeting someone—but not who.

\textbf{Obligation:} 1

\textbf{Backlash:} On a partial, the forgetting is incomplete. The witness remembers a single distinctive detail (a scar, an earring, the color of your eyes). The GM gains 1 SB.

\subsubsection{Rite of the Whispered Step (4 XP)}

\textbf{Tags:} [STEALTH], [MOVEMENT], [SILENCE]

\textbf{Threshold:} A single feather, a pinch of ash, or a moment of absolute stillness.

Your footsteps make no sound for one scene. You may move through dry leaves, broken glass, or loose stone without noise. You may still be seen—but not heard.

\textbf{Obligation:} 1

\textbf{Backlash:} On a miss, your footsteps echo. Mark 1 Fatigue. Any creature within Near range gains +1 die to detect you for the scene.

\subsection{Standard Rites (7-9 XP)}

\subsubsection{Rite of the Spider's Loom (8 XP)}

\textbf{Tags:} [BIND], [AREA], [CONTROL]

\textbf{Threshold:} A spider's web (real or woven from thread) and a drop of your blood.

Weave an invisible web across a zone (Near radius). Any creature entering or leaving the zone must test Body + Athletics (DV 4) or become entangled (cannot move for one exchange). The web lasts for the scene or until burned.

\textbf{Obligation:} 2

\textbf{Backlash:} On a partial, the web catches you as well. You are entangled. The GM gains 1 SB.

\subsubsection{Rite of the Mask of Many Faces (9 XP)}

\textbf{Tags:} [TRANSFORM], [DECEPTION], [VEIL]

\textbf{Threshold:} A mask carved from living wood or woven from silk, and a lock of hair from the person you wish to resemble.

Assume the appearance of any humanoid you have touched. For one scene, your face, voice, and build change to match the target. You do not gain their memories or skills—only their face.

\textbf{Obligation:} 2

\textbf{Backlash:} On a partial, the mask sticks. You cannot remove it for one hour. Mark 1 Fatigue. The GM gains 1 SB.

\subsubsection{Rite of the Unspoken Question (7 XP)}

\textbf{Tags:} [TRUTH], [MIND], [SECRET]

\textbf{Threshold:} A silver mirror or still water, and a question whispered into the surface.

Force a target within Near range to answer one question truthfully. They resist with Resolve (DV 4). If they fail, they must answer—though the answer may be a single word, and they may speak in riddles.

\textbf{Obligation:} 2

\textbf{Backlash:} On a miss, the question turns on you. You must answer one question from the target truthfully. The GM gains 1 SB.

\subsubsection{Rite of the Memory Thread (8 XP)}

\textbf{Tags:} [MEMORY], [ECHO], [SECRET]

\textbf{Threshold:} A grey cord tied around your wrist, and a personal token of the memory you wish to recall.

Recall a single forgotten memory in perfect detail. The memory can be your own or someone else's—if someone else's, you must have a token (hair, blood, a written name). The memory plays before your eyes as if you were there.

\textbf{Obligation:} 2

\textbf{Backlash:} On a partial, the memory is incomplete. You see only fragments. The GM gains 1 SB, and you gain a Condition: \textit{Memory-Frayed} (-1 die to all rolls involving recall) until you rest.

\subsection{High Rites (12-14 XP)}

\subsubsection{Rite of the Silk Grave (12 XP)}

\textbf{Tags:} [DEATH], [BIND], [SECRET]

\textbf{Threshold:} A cocoon of silk (real or woven), and a dying creature or a fresh corpse.

Wrap the target in silk. They are suspended between life and death. They cannot be healed, cannot be harmed, and cannot be questioned. They are simply... paused. The silk lasts for one day, or until cut. If cut, the target returns to their previous state (dead or dying).

\textbf{Obligation:} 3

\textbf{Backlash:} On a miss, the silk wraps around you instead. You are suspended for one hour. The GM gains 2 SB.

\subsubsection{Rite of the Unwritten Vow (13 XP)}

\textbf{Tags:} [VOW], [FATE], [BOND]

\textbf{Threshold:} A blank page, a drop of blood, and a witness.

Swear an oath that is never spoken aloud. The oath is witnessed by the Ghe'hai's patron—Inaea, Ikasha, or the Silent Choir. Breaking the oath causes the oath-breaker to suffer -2 dice to all rolls until they atone. The oath cannot be broken by magical means.

\textbf{Obligation:} 3

\textbf{Backlash:} On a miss, the oath binds you to a term you did not intend. The GM chooses an additional clause. Mark 1 Fatigue.

\subsubsection{Rite of the Ghe'hai's Escape (14 XP)}

\textbf{Tags:} [TELEPORT], [STEALTH], [ESCAPE]

\textbf{Threshold:} A silk cord tied to a stationary object, and a clear path of retreat.

When you are cornered or capture is inevitable, you may vanish. You reappear at a location of your choice within Far range that you can see or have visited within the hour. You leave behind a single silk thread—and a single word (a name, a warning, a promise) written in the air.

\textbf{Obligation:} 3

\textbf{Backlash:} On a partial, you vanish—but something follows. A spirit, a memory, or a Oath-Keeper appears at your destination. The GM gains 2 SB.

\subsection{Shared Ghe'hai Talents}

\subsubsection{Silk-Sense (4 XP)}

\textbf{Requirement:} Ghe'hai training

You can sense any thread, cord, or web within Near range, even if invisible. You know its tension, its anchor points, and roughly what it is connected to. This is a passive ability.

\subsubsection{The Unseen Blade (6 XP)}

\textbf{Requirement:} Ghe'hai training, Melee 2+

When you attack from a hidden position or from a crowd, your target suffers -1 die to defend. They do not see the threat until it is too late.

\subsubsection{Mask-Born (8 XP)}

\textbf{Requirement:} Ghe'hai training, Tier II+

You may spend a downtime scene to create a new identity—a face, a name, a history. The identity holds up to mundane scrutiny. Magical detection requires a contested roll (Spirit + Lore vs. your Presence + Deception). You may maintain up to three identities at once.

\subsection{Roleplaying a Ghe'hai Operative}

\begin{itemize}
\item You do not fight fair. You fight unseen, unheard, unforgettable only in the damage you leave behind.
\item You may be called to infiltrate, to negotiate, to eliminate—or to simply observe. The silk does not distinguish between a blade and a whisper.
\item You may struggle with the cost of masks. You have worn so many faces that you are no longer sure which one is yours.
\item You may be asked to choose—between mission and mercy, between the web and the knife. The Ghe'hai do not have commanders. They have contracts.
\end{itemize}

\subsection{Sample Rite Chain: The Quiet Exit}

Some Ghe'hai Rites are not used alone. They are woven together—a chain of silk and shadow.

\begin{itemize}
\item \textbf{Step 1:} \textit{Rite of the Whispered Step} (Low) — Silence your footsteps.
\item \textbf{Step 2:} \textit{Rite of the Forgotten Face} (Low) — Blur your features.
\item \textbf{Step 3:} \textit{Rite of the Unspoken Question} (Standard) — Learn the patrol schedule from a sleepy guard.
\item \textbf{Step 4:} \textit{Rite of the Ghe'hai's Escape} (High) — Vanish when the alarm is raised.
\end{itemize}

Each Rite in the chain may be cast in sequence without additional actions, but the Obligation cost accumulates. The GM gains 1 SB per Rite beyond the first (the web notices the pattern).

\subsection{GM Guidance: Intrusions for Ghe'hai}

\begin{itemize}
\item A spider lowers itself onto the Ghe'hai's shoulder. In its web is a single perfect word—the name of someone who needs their help.
\item The Ghe'hai's mask begins to itch. The face underneath is not theirs—not anymore.
\item A thread the Ghe'hai tied earlier is now impossible to undo without cutting. Cutting it severs a minor bond—they lose 1 Boon.
\item The Ghe'hai dreams of a loom weaving silk. The pattern is their life. A single thread is about to snap.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
