% --- Patron: Inaea, Angel of the Spider (Webs & Fate) ---
\subsection{Inaea, Angel of the Spider (Webs \& Fate)}
\textit{Lore.} Where Isoka sheds, Inaea binds—threads of debt, favor, and inevitability.

\begin{quote}
Tie one knot for what you owe, two for what you’re owed, and a third for what will answer both.
\end{quote}

\paragraph{Tie a Simple Knot (Low, 4 XP)} \emph{Action; Near; Yes (link once).}
\textbf{Materials:} A single thread.\\
\textbf{Effect:} Declare two minor events linked; either \textbf{force 1 SB} (GM suit) on a foe when the first triggers \emph{or} bank \textbf{+1 die} for a follow-on roll this scene.\\
\textbf{Push It:} The held +1 ignores one minor disruption; the web may also tug an unintended party once.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Tangled Thread (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Tug a web or net.\\
\textbf{Effect:} Invisible snare in a lane/door. First to cross suffers \(-1\) die on next action.\\
\textbf{Push It:} Brief bind (one beat) enabling an ally setup; affects all who cross.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Weave the Strand of Inevitability (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Three colored threads woven.\\
\textbf{Effect:} Link two actors/actions: when A moves, B is exposed. Choose: \textbf{force 1 SB on B} next action \emph{or} \textbf{+2 dice} to one prediction/setup keyed to the link.\\
\textbf{Push It:} Invert once (B cues A). Breaking the link’s fiction creates \textbf{1 SB (Hearts/Clubs)}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Weaver's Glance (Standard, 7 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Watch a spider finish one radial line.\\
\textbf{Effect:} Ask one precise question about in-scene ties; then gain \textbf{+1 effect} on one leverage/pressure action exploiting it.\\
\textbf{Push It:} Surface a hidden tie (Keeper reveals a quiet obligation/fear); mark \emph{Exposure +1}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Bind the Bargain \textnormal{[OATH]} (High, 11 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Silk loop tied around two thumbs, then cut/knotted.\\
\textbf{Effect:} Bind up to two consenting parties to a clear term. Breach \emph{forces 2 SB} on the breaker and leaves a subtle tell until amends.\\
\textbf{Push It:} Widen to a small circle (up to four); each party names a narrow loophole (Keeper approves). Exploiting it generates \textbf{1 SB (Diamonds)}.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.
