% =========================
% PATRON — Isoka, Angel of Serpents
% =========================
\subsection{Isoka, Angel of Serpents (Change \& Shedding)}

\paragraph{Lore.} 
Isoka teaches that every self is temporary. Serpents shed their skin not in weakness but in renewal, leaving the brittle husk behind as proof that transformation is survival. Followers of Isoka learn to embrace disguise, deception, and metamorphosis, casting aside the past as easily as a garment.

\paragraph{Quote.}
\emph{“Do not mourn the skin you shed. It was never meant to last.” — Isoka, Angel of Serpents}

\paragraph{Loosen the Old Skin (Low, 4 XP)} \emph{Scene; Self; Yes (resist only).}
\textbf{Materials:} A discarded snakeskin or a loose thread. \\
\textbf{Effect:} +1 die to resist an ongoing \emph{Condition} this scene \emph{or} re-roll one \textbf{1} on an escape/evasion. \\
\textbf{Push It:} Also ignore one minor movement penalty; you leave behind a token of your old self that others can leverage. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Subtle Shift (Low, 5 XP)} \emph{Scene; Self; No (stable).}
\textbf{Materials:} Palming a small object from one pocket to another. \\
\textbf{Effect:} Fluid demeanor: +1 die to \textbf{Deceive} to pass as a nearby class/profession \emph{or} +1 Effect to blend into a new crowd/site. \\
\textbf{Push It:} Bypass one minor identity check; you must maintain the false role until scene end. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Shed the Former Self (Standard, 8 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Full change of clothing and an adopted mannerism. \\
\textbf{Effect:} +2 dice to resist one named ongoing \emph{Condition}; once/session declare a minor physical contingency retroactively. \\
\textbf{Push It:} Clear a \emph{temporary, identity-based} Minor Condition; your former identity becomes active in the fiction. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Forked Tongue (Standard, 7 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A harmless lie told to a mirror. \\
\textbf{Effect:} Ambiguous persuasion: when you \textbf{Sway} or \textbf{Command}, a success may generate \emph{Diamonds} (leverage) instead of SB. \\
\textbf{Push It:} One carefully worded lie this scene is accepted as truth; the displaced truth seeks return, complicating matters. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Complete Metamorphosis (High, 12 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A complete identity kit (garb, voice, tokens). \\
\textbf{Effect:} Full appearance/voice change; begin \emph{Controlled} on \textbf{Deceive/Stealth}; once/scene declare a minor contingency retroactively. \\
\textbf{Push It:} Spoof scent/biometric for one check; your original identity partially unmoors and acts independently. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Cast-Off History (High, 13 XP)} \emph{Extended; Self; No.}
\textbf{Materials:} Burning or defacing all mundane records of your old life. \\
\textbf{Effect:} On completion, common records/memories of that identity become unreliable. Trackers via that identity suffer $-2$ dice. (Does not foil magic or intimates.) \\
\textbf{Push It:} A plausible “death” is created for the old identity; one intimate senses deception. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.
