\section{The Philosophy of Physical Obligation}

Every Runekeeper owes. That is the first truth of the covenant. The Patron provides power; the Runekeeper provides service. The ledger balances, or it does not.

But a Runic Warrior owes differently.

A scholar writes their debt in ink. Each Rite inscribed, each Obligation recorded, each witness named. The Codex is a book. The debt is a number. When the number grows too large, the scholar pays in coin, in favors, in whispered secrets. The body remains untouched.

A Runic Warrior has no such luxury.

Their debt is written in flesh. A scar is a payment made. A limp is interest accrued. A callus is a promise kept so often that the skin no longer remembers how to be soft.

\subsection{The Scar as Receipt}

When a Runic Warrior invokes a Rite, they do not turn a page. They touch a scar. They press a rune carved into their gauntlet. They trace the knot sewn into their flesh. The Rite lives in the body, and the body remembers the cost.

This is not symbolism. It is mechanics.

A Runic Warrior's Codex is physical. A gorget, engraved with vows. A staff, carved with seasonal runes. A shield, marked with the names of thresholds guarded. The Codex can be damaged. It can be destroyed. But it cannot be stolen in the same way a book can be stolen, because it is worn, carried, or held.

And when the Codex is part of the body—tattoos, brands, scarification—it cannot be stolen at all.

\subsection{The Weight of Visibility}

A traditional Runekeeper can hide. Their Codex fits in a satchel. Their Thiasos can be dismissed or concealed. They walk through a city gate without announcing their covenant.

A Runic Warrior cannot.

The engraved gorget is visible. The rune-staff is a weapon. The tattoos curl up the neck and across the knuckles. Even the sentient blade, quiet as it may be, is still a blade.

This visibility is a cost. The Chain-Lanterns notice. The Cord assesses. The Inquisition takes notes. But it is also a privilege. A Runic Warrior announces their covenant with every step. They do not need to produce a writ. Their body is the writ.

\subsection{Fatigue as Resource, Not Penalty}

In the core rules, Fatigue is a penalty. It accumulates. It worsens Position. It leads to Harm.

For a Runic Warrior, Fatigue is also a resource.

Every scar is a memory of Fatigue endured. Every callus is a history of labor. The Runic Warrior learns to read their own exhaustion as a map—here is where the Rite cost more than expected, here is where the battle turned, here is where the Patron asked for interest and the Runic Warrior paid.

This expansion introduces talents that convert Fatigue into advantage. Not by ignoring it, but by \textit{using} it. A warrior who knows how to fall knows how to rise. A body that has been pushed to its limit knows exactly where the next push will land.

\subsection{The Witness of Wounds}

A Runic Warrior's scars are witnessed.

Not by a notary. Not by a clerk. By anyone who looks.

A scar on the face tells a story. A missing finger asks a question. A brand on the back of the hand declares allegiance.

This is the Runic Warrior's vulnerability. They cannot hide their history. They cannot claim neutrality. Their body is a public ledger, written in tissue and bone.

But it is also their strength. A scar is harder to forge than a seal. A brand is harder to counterfeit than a signature. The Runic Warrior's covenant is visible, undeniable, and carried into every threshold.

\subsection{The Debt That Cannot Be Deferred}

A scholar can defer a debt. They can plead ignorance, claim a misreading, request an extension. The ledger is patient. The ink can be erased.

A Runic Warrior cannot defer.

The scar is already cut. The finger is already gone. The limp does not improve with negotiation.

This is the final philosophy of physical obligation: some debts are paid in advance. The Runic Warrior does not owe a Rite; they \textit{are} the Rite. The covenant is not recorded; it is embodied.

And when the Patron calls, the Runic Warrior does not reach for a book.

They reach for a scar.

They touch the engraved gorget.

They stand.

The thread held. The water carried. That is not a Rite. That is grace.
