\documentclass[10pt,landscape]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{geometry}
\geometry{margin=0.5in, headheight=0pt, footskip=0pt}
\usepackage{multicol}
\usepackage{parskip}
\usepackage{tabularx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{xcolor}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5ex}
\pagestyle{empty}
\definecolor{rulecolor}{RGB}{80,80,80}
\setlength{\columnseprule}{0.4pt}
\setlength{\columnsep}{1em}

\begin{document}
\input{titlepage}
\input{copyright}

\begin{center}
{\huge \textbf{Fate's Edge --- Player's Cyclopedia}} \\[0.2em]
\small \textit{The road remembers. Every choice carries weight.}
\end{center}
\vspace{-0.5ex}
\hrule
\vspace{1ex}
\begin{multicols}{2}

\subsection*{1. The Core Loop}
When you attempt a risky action:
\begin{enumerate}
    \item \textbf{Declare} intent \& approach (Attribute + Skill).
    \item \textbf{GM sets} Difficulty Value (DV) and Position.
    \item \textbf{Roll} dice pool: Attribute + Skill d10s.
    \item \textbf{Count Successes}: each die $\ge 6$ = 1 success. Each die $=1$ = 1 Story Beat (SB) for the GM.
    \item \textbf{Compare} successes to DV $\rightarrow$ use Outcome Matrix.
    \item \textbf{Apply} outcome; GM spends SB for complications.
\end{enumerate}

\subsection*{2. Difficulty Value (DV)}
\begin{tabularx}{\linewidth}{@{} l X @{}}
\textbf{DV 2} & Routine, almost guaranteed \\
\textbf{DV 3} & Pressured (default), mild opposition \\
\textbf{DV 4} & Hard, serious resistance \\
\textbf{DV 5+} & Extreme, dramatic gamble \\
\end{tabularx}

\subsection*{3. Position (Risk)}
\begin{tabularx}{\linewidth}{@{} l X @{}}
\textbf{Dominant}   & Re-roll one failure \\
\textbf{Controlled} & No re-rolls \\
\textbf{Desperate}  & Re-roll one success \\
\end{tabularx}
\vspace{0.5ex}
You may spend \textbf{1 Boon} to improve Position by one step before rolling.

\subsection*{4. Outcome Matrix}
\begin{tabularx}{\linewidth}{@{} l X @{}}
\textbf{Clean Success}      & successes $\ge$ DV, SB = 0: intent achieved, no complication. \\
\textbf{Success with SB}    & successes $\ge$ DV, SB $>$ 0: intent achieved; GM spends SB for a twist. \\
\textbf{Partial}            & $0 <$ successes $<$ DV: progress made; you gain 1 Boon. \\
\textbf{Miss}               & successes = 0: no progress; you gain 2 Boons; GM escalates. \\
\end{tabularx}

\subsection*{5. Boons (Player Currency)}
\begin{itemize}
    \item \textbf{Hold up to 5.} At end of scene, reduce to 2 (excess lost).
    \item \textbf{Earn:} Partial (1 Boon), Miss (2 Boons), bond-driven action, GM award.
    \item \textbf{Spend 1 Boon to:}
    \begin{itemize}
        \item Re-roll a single die after seeing the pool.
        \item Improve Position by one step before rolling.
        \item Activate an on-screen Asset.
        \item Convert 2 Boons $\rightarrow$ 1 XP (once per session, max 2 XP).
    \end{itemize}
\end{itemize}

\subsection*{6. Story Beats (GM Currency)}
Each die $=1$ you roll gives the GM 1 SB. GM spends SB to complicate the scene:
\begin{tabularx}{\linewidth}{@{} l X @{}}
\textbf{1 SB}  & Minor: noise, tick a timer, small setback. \\
\textbf{2 SB}  & Moderate: alarm raised, lose advantage, worsen Position. \\
\textbf{3+ SB} & Major: reinforcements, scene shift, break an asset. \\
\end{tabularx}

\subsection*{7. Fatigue \& Harm}
\textbf{Fatigue:} Each point worsens your Position (Dominant $\rightarrow$ Controlled $\rightarrow$ Desperate). If already Desperate, each Fatigue imposes $-1$ die.
\begin{itemize}
    \item Fatigue track size = Body. When track fills $\rightarrow$ increase Harm by 1, clear all Fatigue.
    \item Short rest (1 hour) removes 2 Fatigue. Full night's rest removes all.
\end{itemize}
\textbf{Harm:}
\begin{tabularx}{\linewidth}{@{} l X @{}}
\textbf{Harm 1} & $-1$ die on related actions \\
\textbf{Harm 2} & $-1$ die on most actions \\
\textbf{Harm 3} & Incapacitated or dying \\
\end{tabularx}
\vspace{0.5ex}
Taking Harm clears all existing Fatigue (the ``roller-coaster'' effect).

\subsection*{8. Combat Quick Reference}
\begin{itemize}
    \item \textbf{Turn:} 1 Action + 1 Move. Move shifts one range band (Close $\leftrightarrow$ Near $\leftrightarrow$ Far).
    \item \textbf{Opportunity Attack:} Leaving Close range without Disengage provokes one free melee attack.
    \item \textbf{Sidestrike:} Move one band as part of a melee attack; worsen Position by one step.
    \item \textbf{Shove:} Opposed Body+Athletics; push target one band.
    \item \textbf{Grapple:} Opposed Body+Melee; both Engaged; target suffers $-1$ die.
\end{itemize}
\vfill\null
\columnbreak

\subsection*{9. Essential Skills (Abridged)}
\textbf{Combat:} Melee, Ranged, Brawl, Tactics. \\
\textbf{Physical:} Athletics, Stealth, Endurance, Craft, Survival. \\
\textbf{Social:} Sway, Command, Deception, Performance, Insight. \\
\textbf{Knowledge:} Lore, Investigation, Medicine. \\
\textbf{Magic:} Arcana, Ritual.

\subsection*{10. Making a Roll -- Example}
Dice pool = Attribute + Skill. \\
Success on $6+$, SB on $1$. \\
\emph{Example:} Body 3 + Melee 2 = 5d10. Roll: 6,4,3,1,10 $\rightarrow$ successes: 6,10 (10 counts as two) = 3 successes. SB: one 1 $\rightarrow$ 1 SB.

\subsection*{11. Quick DC \& Effect Guidelines}
\begin{tabularx}{\linewidth}{@{} l X @{}}
\textbf{DV 2} & Routine, almost guaranteed. \\
\textbf{DV 3} & Standard pressured action. \\
\textbf{DV 4} & Hard, serious opposition. \\
\textbf{DV 5+} & Extreme, dramatic gamble. \\
\end{tabularx}
\vspace{0.5ex}
\textbf{Effect (scope of success):} Limited (minor), Standard (expected), Great (major). GM sets Effect based on fiction.

\subsection*{12. Critical Success (10s)}
A roll of 10 counts as two successes. If the overall roll succeeds:
\begin{itemize}
    \item \textbf{One 10:} Strong success -- improve Position on next action or gain minor advantage.
    \item \textbf{Two 10s:} Exceptional -- choose two benefits.
    \item \textbf{Three 10s:} Legendary -- clear 1 segment from a relevant timer.
\end{itemize}
A 10 is never re-rolled, even by Position effects.

\subsection*{13. Boons vs. XP -- A Reminder}
Boons are dramatic fuel, not XP tokens. Spend them freely to re-roll, improve Position, or activate Assets. Hoarding Boons for conversion (2 Boons $\rightarrow$ 1 XP) slows the table's momentum. The real reward is agency, not a trickle of XP.

\subsection*{14. Golden Rule}
When in doubt, make the choice that serves the story. If you cannot remember a rule, default to \textbf{DV 3, Controlled Position}, and let the dice fall.

\end{multicols}
\vspace{1ex}
\hrule
\begin{center}
\small \textcopyright\ Fate's Edge. This reference may be printed for personal use. \\
\emph{``The coin that never spends is the one you don't remember taking.'' --- Serafine of the Velvet Touch}
\end{center}
\end{document}