\documentclass[12pt,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathpazo}
\usepackage{lettrine}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{epigraph}
\usepackage{verse}
\usepackage{multicol}
\usepackage{titlesec}
\usepackage[margin=1in]{geometry}
\usepackage{afterpage}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}

% Page styling
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[LO]{\leftmark}
\fancyhead[RE]{\rightmark}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% Chapter styling
\titleformat{\chapter}[display]
{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}
\titlespacing*{\chapter}{0pt}{50pt}{40pt}

% Custom commands
\newcommand{\ritualname}[1]{\vspace{10pt}\noindent\textbf{\large #1}\vspace{5pt}}
\newcommand{\shadow}[1]{\textit{#1}}
\newcommand{\whisper}[1]{\textbf{#1}}

% PRODUCT TITLE
% Use on cover, title page, and store listings
\newcommand{\ExpansionTitle}{Fate's Edge: The Whispered Grimoire}
\newcommand{\ExpansionSubtitle}{Rituals \& Rites}
% Requires \usepackage{hyperref}
% SHORT INTERNAL HANDLE (headers, refs)
\newcommand{\ShortRunekeeper}{The Whispered Grimoire}

\begin{document}

\frontmatter

% Title Page
\begin{titlepage}
\centering
\vspace*{2cm}
{\Huge\bfseries Fate's Edge: The Whispered Grimoire\par}
\vspace{1cm}
{\Large A Collection of Rituals, Rites and Forbidden Arts\par}
\vspace{2cm}
{\Large Compiled by Veyla the Hedge-Witch\par}
\vspace{0.5cm}
{\large Servant of Ikasha, She Who Sleeps Within Shadow\par}
\vspace{1cm}
{\large With Notes from Rukhir, My Raven Familiar\par}
\vspace{3cm}
{\large Tenth Moon of the Crimson Year\par}
\vspace{0.5cm}
{\large From the Press of Whispered Pages, Theona\par}
\end{titlepage}

% Copyright Page
\newpage
\thispagestyle{empty}
\vspace*{5cm}
\begin{center}
{\small
The Whispered Grimoire\\
A Collection of Shadow-Rituals and Forbidden Arts\\
\vspace{0.5cm}
Compiled by Veyla the Hedge-Witch\\
\vspace{1cm}
Copyright \copyright\ 2025 Veyla of the Tulkani\\
All rights reserved. No part of this work may be reproduced or transmitted in any form or by any means without written permission from the author.\\
\vspace{1cm}
This work contains dangerous knowledge. Read at your own risk.\\
The author and publisher are not responsible for any consequences arising from the use of these rituals.\\
\vspace{1cm}
Printed in the Three Greens\\
Press of Whispered Pages, Theona\\
Tenth Moon of the Crimson Year
}
\end{center}

% Dedication
\newpage
\thispagestyle{empty}
\vspace*{8cm}
\begin{center}
\begin{verse}
\textit{To Rukhir, my shadow-brother,}\\
\textit{whose black wings carry secrets} \\
\textit{and whose beak pecks at truth} \\
\textit{until it bleeds revelation.} \\
\ \\
\textit{To Ikasha, Mistress of Whispers,}\\
\textit{who teaches us that in shadow} \\
\textit{we find the spaces between} \\
\textit{where power truly dwells.}
\end{verse}
\end{center}

% Table of Contents
\tableofcontents

\mainmatter

\chapter{To the Seeker of Shadows}

\lettrine[lines=3]{L}{}et me begin, dear reader, with a warning whispered in the dark hours before dawn, when the veil between worlds grows thin as a spider's silk. What you hold in your hands is not mere parchment and ink, but a gathering of truths that the light-bound would burn, the law-bound would imprison, and the truth-bound would deny.

I am Veyla, hedge-witch of the Tulkani, daughter of wandering merchants who taught me that the greatest treasures are those hidden in plain sight. My path led me to the worship of Ikasha, She Who Sleeps Within Shadow, and through her dark blessings, I have learned that power flows not from the sun's harsh light, but from the spaces between the stars.

\shadow{Rukhir tells me you are worthy of these words. He has been watching you, as ravens watch all who walk the paths between. "She moves with the caution of one who knows that shadows have teeth," he caws to me. "Her footsteps make no sound on the stones of secrecy." For a raven to speak of caution is high praise indeed.}

The rituals contained within these pages are not for the faint of heart or the weak of spirit. Each one demands a price, for Ikasha's gifts are never given freely. But for those with the wit to pay and the will to claim, they offer power that kings would kill for and gods would envy.

\section*{The Nature of Shadow-Magic}

Shadow is not darkness, child of light. It is the space between what is and what could be, the pause between heartbeats where truth whispers its name. In my years of walking the twilight paths, I have learned that shadow-magic is not about casting spells in the traditional sense, but about opening doors that others have locked and forgotten.

\shadow{Rukhir nods from his perch upon my shoulder. "The simple think shadow is absence," he says, preening his black feathers. "The wise know it is presence waiting to be acknowledged."}

The Tulkani understand this better than most peoples. We are wanderers, yes, but we are also watchers. In our travels, we have learned that every culture hides its secrets in the spaces between words, and it is in those spaces that true power lies. The shadow-magics I present here are drawn from a dozen traditions, whispered to me by spirits, traded with other seekers, and discovered through my own communion with Ikasha.

\section*{The Price of Power}

Before we begin, you must understand a fundamental truth: there is no such thing as free magic. The Utarans speak of balance, the Aeler of the deep laws, and the Wood-Elves of the cycle of giving and taking. In the tongue of my people, we say that every gift must be paid for in kind, and Ikasha's gifts are the most expensive of all.

\whisper{Listen well to this, for it is the most important lesson I can teach you:}

\begin{verse}
The shadow that you command \\
Will command you in return. \\
The secret that you steal \\
Will steal a piece of you. \\
The truth that you uncover \\
Will uncover truths about yourself \\
That you may not wish to know.
\end{verse}

\shadow{Rukhir caws agreement. "I have seen wizards become slaves to their own spells," he tells me. "I have watched sorceresses consumed by the very shadows they thought to control."}

This is why I have included not just the rituals themselves, but the warnings, the preparations, and the aftermaths. Each ritual in this grimoire comes with three sections: The Weaving (how to perform the ritual), The Price (what you must give), and The Whisper (what Ikasha will whisper to you in return).

\section*{The Tools of the Shadow-Witch}

You will find that shadow-magic requires few material components compared to the flashy spells of the light-wielders. Our power comes from understanding rather than force, from knowing rather than doing. However, there are certain tools that every shadow-witch should possess:

\begin{itemize}
\item \textbf{A black candle} - Not for light, but for focus. The flame should gutter and dance as if in a wind that others cannot feel.
\item \textbf{A mirror of dark glass} - To see what others cannot see, and to guard against what would see you.
\item \textbf{A raven's feather} - For writing in shadow-ink that only the initiated can read.
\item \textbf{A small pouch of shadow-salt} - Gathered at the crossroads when the moon is new, used to mark boundaries and ward off unwanted attention.
\item \textbf{A whisper-stone} - A smooth stone that has been held in the mouth during a silence, used to store and release spoken secrets.
\end{itemize}

\shadow{Rukhir has brought me a bone from a hanged man, which he claims is essential for certain workings. I have not yet had the courage to use it.}

\section*{The Structure of This Work}

This grimoire is divided into four sections, each corresponding to a different aspect of shadow-magic:

\begin{enumerate}
\item \textbf{Whispers and Secrets} - Rituals for gathering information, speaking across distances, and uncovering hidden truths.
\item \textbf{Shadows and Veils} - Rituals for concealment, transformation, and movement through spaces that should not exist.
\item \textbf{Dreams and Nightmares} - Rituals for entering the dream-realm, influencing sleepers, and drawing power from the unconscious mind.
\item \textbf{Bargains and Bindings} - Rituals for making pacts with spirits, binding servants, and compelling others to your will.
\end{enumerate}

Each ritual is presented in the same format:

\ritualname{Name of the Ritual}
\textbf{Purpose:} What the ritual accomplishes\\
\textbf{Weaving:} Step-by-step instructions for performance\\
\textbf{Price:} What must be sacrificed or risked\\
\textbf{Whisper:} What Ikasha grants in return\\
\textbf{Warning:} Important cautions and potential consequences

\shadow{Rukhir insists I add this final note: "Not all who seek power are worthy of it, and not all who are worthy seek it. Be certain which you are before you turn the next page."}

Now, if you have read this far and still wish to continue, then perhaps you are ready for what lies ahead. Light your black candle, settle into your chair, and prepare to learn the arts that others fear to name.

But remember, dear seeker - once you have seen into the shadows, you can never claim to walk in the light again.

\chapter{Whispers and Secrets}

\section*{The Art of Listening}

In the Tulkani tradition, we say that the greatest magic is not in speaking, but in hearing what others dare not say. The rituals in this chapter will teach you to listen to the whispers that exist in the spaces between words, to read the truth in a person's eyes, and to gather secrets that would make emperors tremble.

\shadow{Rukhir has been very quiet this evening. I think he senses that we are approaching dangerous territory. Even ravens have their secrets, it seems.}

\ritualname{The Listening Shadow}
\textbf{Purpose:} To hear conversations from great distances, particularly those that the speakers believe are private.

\textbf{Weaving:} 
\begin{enumerate}
\item Light a black candle at the darkest hour of night, when even the moon has set.
\item Take a raven's feather and whisper into it the name of the person whose words you wish to hear.
\item Bury the feather beneath your doorstep, or in the earth beneath a window where your target might pass.
\item For three nights, you must dream of shadows dancing. Do not fight these dreams - embrace them.
\item On the fourth night, the words you seek will come to you on the wind, spoken in a voice like rustling leaves.
\end{enumerate}

\textbf{Price:} You will hear one conversation that you desperately wish not to hear. This knowledge will burden you until you either act upon it or deliberately forget it.

\textbf{Whisper:} Ikasha will grant you the ability to distinguish truth from lies in any spoken word, but only for as long as you hold your breath.

\textbf{Warning:} Do not use this ritual to spy upon those who are innocent of wrongdoing. Ikasha favors those who seek justice, not those who indulge in cruelty.

\shadow{Rukhir shifts uncomfortably. "I have seen this ritual drive people to madness," he mutters. "The knowledge comes with the weight of responsibility."}

\section*{The Bonding Ceremony}
\textit{A ritual to strengthen connections between companions}

\textbf{Purpose:} Create a permanent bond between party members that provides mechanical and narrative benefits

\textbf{Participants:} All willing party members (3-5 helpers maximum)

\textbf{Channel Phase (Wits + Bonds):}
\begin{enumerate}
\item Each participant holds an object representing their connection to the group
\item Perform a shared activity (sharing a meal, telling stories, etc.) while focusing on the bond
\item Roll Wits + Bonds skill for each participant
\end{enumerate}

\textbf{Weave Phase (Presence + Sway):}
\begin{enumerate}
\item Participants speak words of commitment to each other
\item Create a physical symbol of the bond (knotting cords, mixing earth from homelands, etc.)
\item Roll Presence + Sway for the primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} Temporary tension between bonded members (-1 die to cooperative actions for Significant Time)
\item \textbf{3-4 SB:} The bond creates unwanted attention from external forces; party generates 2 SB whenever they act as a group for the next session
\item \textbf{5+ SB:} All existing bonds between party members are temporarily severed; party suffers -2 dice to all social rolls for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Bonded members gain +1 die when aiding each other for one arc
\item Can sense each other's general emotional state across any distance for one arc
\item One "bond-mate" can be called upon for aid once per session (counts as Cap 3 follower for that action)
\end{itemize}

\shadow{Rukhir preens thoughtfully. "I have seen flocks torn apart by too-strong bonds," he caws. "The heart that beats as one can also break as one."}

\textbf{Duration:} One full arc, renewable with simpler weekly reaffirmation

\textbf{Casting Time:} Significant Time (ritual requires shared meal and ceremony)

\section*{The Sanctuary Consecration}
\textit{A ritual to create a safe haven from supernatural threats}

\textbf{Purpose:} Establish a protected area that repels hostile supernatural entities

\textbf{Primary Caster:} Spirit 4+, Lore 3+

\textbf{Helpers:} 2-3 participants with relevant skills (Lore, Faith, Arcana)

\textbf{Channel Phase (Spirit + Faith/Lore):}
\begin{enumerate}
\item Gather materials sacred to the caster's beliefs or relevant to the threats faced
\item Consecrate the space through prayer, meditation, or ritual preparation
\item Roll Spirit + relevant skill for each participant
\end{enumerate}

\textbf{Weave Phase (Wits + Ritual):}
\begin{enumerate}
\item Perform the consecration ceremony within the designated area
\item Create permanent wards using blessed/sacred materials
\item Roll Wits + Ritual for the primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The sanctuary attracts the attention of other supernatural entities; 2 SB added to next supernatural encounter
\item \textbf{3-4 SB:} The protection creates a vulnerability elsewhere; choose one party member who becomes more susceptible to supernatural influence for one session
\item \textbf{5+ SB:} The ritual backfires, making the area supernaturally dangerous; the space becomes a beacon for hostile entities
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Area becomes sanctified ground (Sanctity +2) against specified supernatural threats for one month
\item Hostile entities of the specified type suffer -2 dice to actions within the area
\item Area provides +1 die to resistance rolls against specified supernatural effects for one month
\item Party members resting within the sanctuary recover 1 additional Fatigue per night
\end{itemize}

\shadow{Rukhir circles the perimeter warily. "Sacred spaces are like nests," he observes. "They keep some things out, but they also keep you in."}

\textbf{Duration:} One month, renewable with weekly maintenance (10 minutes of ritual work)

\textbf{Casting Time:} Significant Time (requires preparation and ceremony)

\section*{The Memory Reconstruction}
\textit{A ritual to recover lost or suppressed memories}

\textbf{Purpose:} Retrieve information from a participant's suppressed or forgotten memories

\textbf{Primary Caster:} Wits 4+, Arcana 3+

\textbf{Helpers:} 1-2 participants with Insight or Lore skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Participant enters a deep meditative state
\item Primary caster prepares scrying tools or ritual components
\item Roll Wits + Arcana for primary caster, Wits + Insight/Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Spirit + Insight):}
\begin{enumerate}
\item Guide the participant through their memories using guided visualization
\item Use ritual tools to stabilize recovered memories
\item Roll Spirit + Insight for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} Recovered memories are fragmented or unreliable; information may be misleading
\item \textbf{3-4 SB:} The ritual triggers traumatic memories causing temporary psychological effects; participant suffers -1 die to social rolls for Significant Time
\item \textbf{5+ SB:} Participant becomes obsessed with recovered memories; must make Spirit + Resolve (DV 2) rolls to act against the memory's implications for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Recover one specific memory or set of memories from the participant's past (immediate)
\item Memory clarity depends on success margin (partial recall vs. complete recovery)
\item Participant gains +2 dice on related knowledge rolls for this scene
\item GM reveals one piece of information that connects the recovered memory to current events
\end{itemize}

\shadow{Rukhir tilts his head. "The past is a cage made of whispers," he says. "Some birds prefer to forget they were ever caged."}

\textbf{Duration:} Memory effects last for this scene; knowledge benefits last for this session

\textbf{Casting Time:} Some Time (requires deep meditation and guided ritual)

\section*{The Shadow Passage}
\textit{A ritual to travel through the Ways Between}

\textbf{Purpose:} Transport the party across great distances through shadow-realm shortcuts

\textbf{Primary Caster:} Wits 5+, Arcana 4+, Spirit 3+

\textbf{Helpers:} All party members (spiritual support during the journey)

\textbf{Channel Phase (Spirit + Arcana):}
\begin{enumerate}
\item Create a circle of shadow-conductive materials (obsidian, black candles, etc.)
\item Each participant contributes a drop of blood or lock of hair to the circle
\item Roll Spirit + Arcana for primary caster, Spirit + Resolve for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Ritual):}
\begin{enumerate}
\item Perform the opening ceremony to breach the shadow-realm
\item Lead the party through the passage while maintaining the connection
\item Roll Wits + Ritual for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The journey takes longer than expected, arriving at an inconvenient time (advance relevant clocks by 1 segment)
\item \textbf{3-4 SB:} The party arrives in a slightly different location than intended; position for first action is Risky instead of Controlled
\item \textbf{5-6 SB:} Something follows the party back from the shadow-realm; introduce a persistent supernatural threat that tracks the party
\item \textbf{7+ SB:} The party finds themselves in the Ways Between.
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Transport party instantly across great distances (anywhere on the same continent) for this scene
\item Arrival position is Controlled for planned destinations, Risky for improvised ones
\item +1 effect on the first action after arrival (momentum from the journey)
\item Party is immune to fatigue from the travel itself
\item Cannot be used more than once per week without increasing backlash severity
\end{itemize}

\shadow{Rukhir fluffs his feathers nervously. "The spaces between spaces have their own hunger," he warns. "They take payment in ways you won't expect until it's too late."}

\textbf{Duration:} Instantaneous travel with lingering effects for this scene

\textbf{Casting Time:} Significant Time (complex ritual requiring careful preparation)

\section*{The Geas of Binding}
\textit{A ritual to create unbreakable oaths or compulsions}

\textbf{Purpose:} Bind a willing or unwilling target to a specific course of action

\textbf{Primary Caster:} Presence 4+, Spirit 4+, Lore 3+

\textbf{Helpers:} 1-2 participants with relevant cultural or religious knowledge

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Prepare the binding materials according to cultural tradition
\item Research the target's nature and vulnerabilities
\item Roll Spirit + Lore for primary caster, relevant skill for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Perform the binding ceremony with the target present
\item Speak the terms of the geas in the appropriate ritual language
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The geas has unintended loopholes or consequences; target finds creative ways to technically fulfill the geas while undermining its intent
\item \textbf{3-4 SB:} The caster becomes partially bound by the same geas; must also follow a related constraint for Significant Time
\item \textbf{5+ SB:} The binding creates a spiritual debt that attracts unwanted supernatural attention; 3 SB added to next supernatural encounter involving the caster
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Target is compelled to follow the terms of the geas or suffer spiritual consequences for one arc
\item Breaking the geas requires a successful Spirit + Resolve roll (DV 4+) each time
\item Failure to follow the geas causes 1 SB per day of violation
\item Geas can be complex but must be clearly defined (no "be good" - specify actions)
\item Willing targets suffer reduced backlash; unwilling targets may resist with opposed roll
\end{itemize}

\shadow{Rukhir caws sharply. "Words are chains made of air," he says. "But they bind tighter than iron when spoken with true power."}

\textbf{Duration:} One full arc, permanent unless specifically countered by equal magic

\textbf{Casting Time:} Significant Time (formal ceremony requiring proper ritual components)

\section*{The Communion of the Ancients}
\textit{A ritual to speak with the spirits of the dead}

\textbf{Purpose:} Contact and communicate with deceased individuals or ancestral spirits

\textbf{Primary Caster:} Spirit 4+, Lore 3+, Faith 2+

\textbf{Helpers:} 1-2 participants with spiritual sensitivity or relevant cultural knowledge

\textbf{Channel Phase (Spirit + Faith):}
\begin{enumerate}
\item Prepare the ritual space according to cultural traditions for contacting the dead
\item Offer appropriate sacrifices or prayers to the spirits
\item Roll Spirit + Faith for primary caster, Spirit + Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Lore):}
\begin{enumerate}
\item Open the channel to the spirit realm
\item Guide the summoned spirit toward communication
\item Roll Wits + Lore for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The contacted spirit is confused or unhelpful; information provided is vague or misleading
\item \textbf{3-4 SB:} Other, less welcome spirits are drawn to the ritual; 2 SB added to next supernatural encounter
\item \textbf{5+ SB:} The ritual creates a permanent spiritual connection that cannot be easily severed; caster suffers -1 die to all social rolls involving living beings for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Communicate with one specific deceased individual or type of ancestral spirit for Some Time
\item Spirit can provide information, advice, or warnings about specified topics
\item Spirit's knowledge is limited by their mortal understanding and perspective
\item Communication lasts for a number of minutes equal to the success margin
\item Spirit may request a favor or service in return for their cooperation
\end{itemize}

\shadow{Rukhir shivers, feathers ruffling. "The dead remember everything," he whispers. "But they forget what it means to be kind to the living."}

\textbf{Duration:} Some Time (ritual communication period)

\textbf{Casting Time:} Significant Time (preparation and ceremony required)

\section*{The Ward of Protection}
\textit{A ritual to create lasting magical defenses}

\textbf{Purpose:} Establish permanent or semi-permanent protective wards

\textbf{Primary Caster:} Wits 4+, Arcana 3+, relevant cultural skill

\textbf{Helpers:} 2-3 participants with Craft, Engineering, or cultural ritual skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Gather materials appropriate to the type of protection desired
\item Prepare the location where wards will be placed
\item Roll Wits + Arcana for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Craft/Engineering):}
\begin{enumerate}
\item Inscribe, construct, or otherwise create the physical components of the wards
\item Activate the wards through ritual consecration
\item Roll Wits + relevant skill for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The wards have a minor vulnerability or require specific activation conditions; wards function at -1 effectiveness until properly calibrated
\item \textbf{3-4 SB:} The wards attract the attention of entities they're designed to repel; 2 SB added to next encounter with specified threat type
\item \textbf{5+ SB:} The wards backfire, creating the opposite effect or affecting unintended targets; wards harm allies instead of protecting them for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create wards that provide +2 dice to defense against specified threats for one month
\item Wards can cover an area, protect specific individuals, or guard objects
\item Duration: One month, renewable with monthly maintenance rituals
\item Wards can be complex (multiple effects) but each additional effect increases backlash risk
\item Wards can be detected by those with appropriate skills and may reveal the caster's presence
\end{itemize}

\shadow{Rukhir examines the warding materials critically. "Boundaries are promises written in power," he observes. "But promises can be broken, and power can be hungry."}

\textbf{Duration:} One month, renewable with weekly maintenance

\textbf{Casting Time:} Significant Time (complex construction and consecration process)

\section*{The Veil of Misdirection}
\textit{A ritual to conceal the party's true nature and intentions}

\textbf{Purpose:} Create illusions and false impressions that mask the party's activities

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Deception 2+

\textbf{Helpers:} 2-3 participants with Performance, Deception, or cultural masking skills

\textbf{Channel Phase (Wits + Deception):}
\begin{enumerate}
\item Prepare illusion components (cosmetics, costumes, symbolic objects)
\item Each participant contributes a personal item that represents their true self
\item Roll Wits + Deception for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Performance):}
\begin{enumerate}
\item Perform the masquerade ritual, weaving false identities around each participant
\item Create shared illusion patterns that reinforce the group's fabricated story
\item Roll Presence + Performance for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The illusions have minor flaws or inconsistencies; -1 die to social rolls when under scrutiny
\item \textbf{3-4 SB:} The false identities begin to influence the participants' behavior; party members must make Wits + Resolve (DV 2) rolls to act against their assumed personas for Significant Time
\item \textbf{5+ SB:} The ritual creates a feedback loop where the false identities become partially real; participants suffer identity confusion and -2 dice to rolls involving their true skills for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Party assumes completely fabricated identities that are believable to casual observers for one session
\item Each participant gains +2 dice to Deception and Performance rolls related to their assumed identity
\item Social interactions start Controlled when the false identities are relevant
\item Magical detection of true identities suffers -2 dice for Significant Time
\item Party can coordinate complex deceptions as a group action with +1 effect
\end{itemize}

\shadow{Rukhir preens his feathers with satisfaction. "I have worn many faces," he crows. "But I always remember which one is mine."}

\textbf{Duration:} One full session, requiring daily renewal (10 minutes of ritual work)

\textbf{Casting Time:} Significant Time (elaborate preparation and group ceremony)

\chapter{Part II: The Shadow's Embrace}

\section*{On Possession and the Compromised Soul}

\lettrine[lines=3]{T}{}he deepest shadows, dear reader, are not those that fall from candlelight or moonbeam. They are the darkness that grows within, fed by our own fears, our own compromises, our own willingness to trade pieces of ourselves for power. 

\shadow{Rukhir has not spoken much of late. I think he senses the weight of what I must record in these pages. Even ravens, I have learned, can be touched by shadows that should remain buried.}

In all my years walking the twilight paths, I have seen many seekers of power become its prisoners. The line between master and servant is thinner than spider's silk, and it snaps more easily than most would believe. When I speak of possession, I do not merely mean the crude taking of bodies by wandering spirits—though that too has its place in the dark arts. No, the possession I warn you of is subtler, more insidious, and far more dangerous.

\textbf{Player Character Possession:} This is not a condition to be taken lightly, nor one that should ever be inflicted without the player's consent and understanding. Possession represents the ultimate compromise of character agency—the point at which external forces have so thoroughly corrupted or influenced a character that they are no longer fully in control of their actions.

The Compromised condition of possession manifests in stages:

\begin{itemize}
\item \textbf{Stage 1 (Influenced):} Character begins to hear whispers, feel urges, or experience compulsions that are not their own. They retain full agency but must resist temptation through appropriate rolls.
\item \textbf{Stage 2 (Controlled):} Character occasionally acts against their will or better judgment. The possessing force can compel specific actions once per session, which the player may resist with a Spirit + Resolve roll.
\item \textbf{Stage 3 (Possessed):} Character becomes a puppet of the possessing entity. The GM takes control of the character for extended periods, though major life-or-death decisions should still involve player consultation.
\end{itemize}

\shadow{"Possession is not defeat," Rukhir finally speaks, his voice unusually grave. "It is transformation. And transformation, my dear Veyla, is always a kind of death."}

\textbf{GM Guidance on Possession:}

Possession should never be used as a punitive measure or to railroad players. It is a narrative tool of last resort, representing the culmination of a character's moral and spiritual decline. Before imposing possession, the GM should:

\begin{enumerate}
\item Ensure the player understands and consents to the possibility
\item Provide multiple opportunities for the character to resist or break free
\item Make the descent into possession a gradual process with clear warning signs
\item Offer meaningful ways for other characters to help their companion
\item Respect the player's agency even when the character is compromised
\end{enumerate}

The path to possession should be paved with choices—each one a small surrender, each one making the next compromise easier. A character who deliberately seeks dark power should understand that they are playing with forces that demand payment in ways they may not anticipate.

\section*{The Ritual of Exorcism}

When possession takes root, when the shadow has grown too deep, there remains one path to salvation: the Ritual of Exorcism. But know this—exorcism is not a simple spell cast and forgotten. It is warfare. It is negotiation. It is sometimes surrender.

\textbf{Purpose:} To forcibly remove a possessing entity from a possessed subject

\textbf{Primary Caster:} Spirit 4+, Faith 3+, Lore 2+

\textbf{Helpers:} 2+ participants with Faith, Spirit, or relevant cultural/religious skills

\textbf{The Exorcism Encounter:}

Exorcism functions as a special encounter with its own clock mechanics. The ritual creates a contested space where the caster's will battles the possessing entity for control of the subject's soul.

\textbf{Exorcism Clock:} 2 + Subject's Tier in segments (minimum 4 segments)

Each segment represents a phase of spiritual combat requiring:
\begin{enumerate}
\item \textbf{Channel Phase (Spirit + Faith):} Primary caster and helpers focus their spiritual energy
\item \textbf{Weave Phase (Presence + Command):} Direct confrontation with the possessing entity
\item \textbf{Response Phase:} The entity fights back—participants are Exhausted and cannot act for One Moment
\end{enumerate}

\textbf{Entity Response Actions:}
During each Response Phase, the possessing entity may attempt actions that, if not resisted, advance the Exorcism Clock:
\begin{itemize}
\item \textbf{Whispers of Doubt:} Tempt participants with promises or threats
\item \textbf{Physical Manifestation:} Cause harm to participants or the subject
\item \textbf{Spiritual Assault:} Drain spiritual energy from participants
\item \textbf{Reality Distortion:} Create illusions or false visions to confuse the ritual
\end{itemize}

Participants may resist these actions with appropriate rolls, but failure allows the clock to advance by one segment.

\textbf{Defensive Support:}
Non-participating allies may concentrate on defense, providing +1 die to resistance rolls for each defender. However, defenders cannot take other actions during the exorcism.

\textbf{Ritual Failure:}
If the Exorcism Clock fills completely before the ritual succeeds, the exorcism fails with severe consequences:
\begin{itemize}
\item The possessing entity grows stronger (+1 die to all future possession attempts)
\item The subject suffers additional corruption (advance Sanity loss or equivalent)
\item All participants suffer Fatigue 2 from spiritual exhaustion
\item The entity gains knowledge of the participants' weaknesses
\end{itemize}

\textbf{For Player Characters:}
When a PC becomes possessed and exorcism fails, they gain complications based on the possessing entity's Tier:
\begin{itemize}
\item \textbf{Tier I-II Entities:} 1 SB banked each session from the entity's influence
\item \textbf{Tier III-IV Entities:} 2 SB banked each session from the entity's influence  
\item \textbf{Tier V+ Entities:} 3+ SB banked each session from the entity's influence
\end{itemize}

These banked SB represent the entity's ongoing influence and may manifest as:
\begin{itemize}
\item Compulsive behaviors or speech patterns
\item Nightmares or visions that provide the entity information
\item Uncontrolled moments where the entity acts through the character
\item Social consequences as others notice the character's changed nature
\end{itemize}

\shadow{Rukhir shivers, and I realize his feathers have gone dull. "Even I," he says quietly, "have felt the touch of things that would wear me like a glove. The difference, Veyla, is that I chose to be worn."}

\textbf{Player Agency Preservation:}
While possessed characters suffer ongoing complications, their players must always retain meaningful agency:
\begin{itemize}
\item Players choose when to resist the entity's influence
\item Major character decisions remain with the player
\item The entity cannot force the character to betray core relationships without resistance
\item Redemption paths should always remain possible, however difficult
\end{itemize}

The Ritual of Exorcism is not merely about casting spells—it is about the fundamental struggle between light and shadow, freedom and control, self and other. Approach it with the gravity it deserves, and remember that sometimes, the greatest mercy is to help a friend find peace, even if that peace requires them to let go of everything they were.

\textbf{Duration:} Variable encounter time, typically 20-30 minutes of game time per segment

\textbf{Casting Time:} Extended ritual requiring continuous focus and participation

\section*{Mist-Road Veil}
\textit{Draws a low fog that dulls footfalls and blurs silhouettes along a chosen path}

\textbf{Purpose:} Create concealing mist that aids stealthy movement along a specific route

\textbf{Primary Caster:} Wits 3+, Arcana 2+, Survival 2+

\textbf{Helpers:} 1-2 participants with Survival or Stealth skills

\textbf{Channel Phase (Wits + Survival):}
\begin{enumerate}
\item Gather water sources and cooling materials (damp cloth, morning dew, etc.)
\item Trace the path where mist is desired with light touch
\item Roll Wits + Survival for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Arcana):}
\begin{enumerate}
\item Channel moisture into the air along the designated path
\item Whisper words of concealment in the local tongue
\item Roll Wits + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} Mist is patchy or reveals more than it conceals; -1 die to Stealth rolls
\item \textbf{3-4 SB:} Mist attracts unwanted attention from creatures that hunt in fog; 2 SB added to next encounter
\item \textbf{5+ SB:} Mist becomes supernaturally thick, disorienting even the caster; party suffers -2 dice to navigation for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Creates concealing fog along a path up to 100 yards long for Significant Time
\item Party members moving along the path gain +2 dice to Stealth rolls
\item Sound is muffled by 50\% within the mist
\item Non-magical tracking becomes nearly impossible within the affected area
\end{itemize}

\shadow{Rukhir circles the area warily. "Mist hides the hunter," he caws, "but it also hides the hunted. Choose your prey wisely."}

\textbf{Duration:} Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (10-15 minutes of preparation and ritual)

\section*{Bell of Truce}
\textit{Rings a hospitality note that compels all within earshot to parley before violence}

\textbf{Purpose:} Create a magical compulsion for peaceful negotiation before combat

\textbf{Primary Caster:} Presence 4+, Spirit 3+, Command 2+

\textbf{Helpers:} 1-2 participants with Sway or Performance skills

\textbf{Channel Phase (Spirit + Command):}
\begin{enumerate}
\item Obtain a bell or chime of significant quality
\item Focus on the concept of hospitality and peaceful resolution
\item Roll Spirit + Command for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Performance):}
\begin{enumerate}
\item Ring the bell with specific intent and rhythm
\item Speak words of truce in the old tongue
\item Roll Presence + Performance for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The compulsion is weak or easily resisted; -1 die to Sway rolls for next social encounter
\item \textbf{3-4 SB:} The bell's influence extends beyond intended area, affecting neutral parties; 2 SB added to crowd-related complications
\item \textbf{5+ SB:} The compulsion backfires, making the caster appear weak or cowardly; -2 dice to Command rolls for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item All intelligent beings within 200 yards must make Spirit + Resolve (DV 3) to initiate violence for Some Time
\item Successful resistance grants +1 die to subsequent social rolls with that individual
\item Failure to parley first creates 2 SB backlash when violence finally erupts
\item The bell's tone marks a sacred space of temporary sanctuary
\end{itemize}

\shadow{Rukhir preens. "Even warriors must pause to hear beautiful sounds," he observes. "It is in that pause that wisdom often speaks."}

\textbf{Duration:} Some Time (approximately 30 minutes)

\textbf{Casting Time:} Some Time (ritual ringing and proclamation)

\section*{Lantern Weir}
\textit{Stitches light into a soft fence that guides crowds and resists panic surges}

\textbf{Purpose:} Create a barrier of light that guides movement and prevents chaotic rushes

\textbf{Primary Caster:} Wits 3+, Arcana 3+, Performance 2+

\textbf{Helpers:} 2-3 participants with Performance or Command skills

\textbf{Channel Phase (Wits + Performance):}
\begin{enumerate}
\item Gather multiple light sources (lanterns, candles, etc.)
\item Arrange participants in a line or curve to form the barrier
\item Roll Wits + Performance for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Weave the light sources together into a unified barrier
\item Speak words of guidance and protection
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The light barrier flickers or has gaps; -1 die to Crowd management rolls
\item \textbf{3-4 SB:} The barrier attracts attention from the wrong crowd; 2 SB added to next social encounter
\item \textbf{5+ SB:} The light becomes blinding rather than guiding; all within 50 yards suffer -1 die to perception rolls
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Creates a 50-foot barrier of soft, guiding light for Significant Time
\item Crowd movement is directed and calmed; +2 dice to Crowd management rolls
\item Panic effects are reduced by 2 SB for those within the barrier
\item The barrier provides dim illumination equivalent to moonlight
\end{itemize}

\shadow{Rukhir examines the light sources critically. "Light is a gentle master," he says, "but it demands respect from both those who wield it and those who would flee from it."}

\textbf{Duration:} Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (15-20 minutes of arrangement and ritual)

\section*{Oath of Shared Breath}
\textit{Binds the party's rhythm; fatigue spreads thinly among all who consent}

\textbf{Purpose:} Distribute fatigue and harm among willing party members

\textbf{Primary Caster:} Presence 4+, Spirit 3+, Bonds 2+

\textbf{Helpers:} All willing party members (2+ participants required)

\textbf{Channel Phase (Spirit + Bonds):}
\begin{enumerate}
\item All participants join hands or touch in a circle
\item Each participant speaks their willingness to share the burden
\item Roll Spirit + Bonds for primary caster, Spirit + Resolve for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Weave the participants' life forces together through ritual words
\item Create a shared rhythm of breath and heartbeat
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The sharing is uneven, favoring some over others; one participant suffers +1 Fatigue
\item \textbf{3-4 SB:} The bond creates unwanted emotional connections; party suffers -1 die to social rolls with outsiders for Significant Time
\item \textbf{5+ SB:} The sharing becomes too intimate; participants can sense each other's thoughts and emotions for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item All willing participants share Fatigue and Harm equally for one session
\item Total Fatigue is divided by number of participants (rounded up)
\item Harm from attacks is distributed among all participants
\item Participants can sense each other's general health and location within 100 yards
\end{itemize}

\shadow{Rukhir tilts his head. "The heart that beats alone," he says, "can stop alone. But hearts that beat together... those are harder to silence."}

\textbf{Duration:} One full session

\textbf{Casting Time:} Some Time (ritual circle and shared oath)

	\section*{River's Favor}
	\textit{Coaxes currents to slacken at a ford or quay so boats and rafts pass cleanly}

	\textbf{Purpose:} Calm water currents to aid safe passage for watercraft

	\textbf{Primary Caster:} Wits 3+, Arcana 2+, Survival 3+

	\textbf{Helpers:} 1-2 participants with Survival or Pilot skills

	\textbf{Channel Phase (Wits + Survival):}
	\begin{enumerate}
	\item Touch the water at the point of desired calm
	\item Speak to the river's spirit or natural forces
	\item Roll Wits + Survival for primary caster, relevant skills for helpers
	\end{enumerate}

	\textbf{Weave Phase (Spirit + Arcana):}
	\begin{enumerate}
	\item Channel calming energy into the water
	\item Create a path of gentle currents
	\item Roll Spirit + Arcana for primary caster
	\end{enumerate}

	\textbf{Backlash:}
	\begin{itemize}
	\item \textbf{1-2 SB:} The calm is temporary or localized; -1 die to Pilot rolls
	\item \textbf{3-4 SB:} The river's natural flow is disrupted elsewhere; 2 SB added to next water-related encounter
	\item \textbf{5+ SB:} The ritual angers local water spirits; introduce a water-based supernatural threat
	\end{itemize}

	\textbf{Effect:}
	\begin{itemize}
	\item Creates a 100-foot stretch of calm water for Significant Time
	\item Watercraft gain +2 dice to Pilot rolls within the affected area
	\item Currents are reduced by 50\% within the calm zone
	\item Small debris and obstacles are pushed to the edges of the calm area
	\end{itemize}

	\shadow{Rukhir drinks from the river cautiously. "Water remembers promises," he says, "but it also remembers betrayals. Be gentle in your asking."}

	\textbf{Duration:} Significant Time (approximately 1 hour)

	\textbf{Casting Time:} Some Time (communion with the water and ritual work)

	\chapter{Rites of the Binding Arts}
	\label{chap:outer-rites}

	\lettrine[lines=3]{O}{}ut there, beyond the lamplight and the last safe mile of road, the world unfolds into angles that do not ask permission to exist. Some call them the outer spheres; I call them the \emph{places that look back}. If you would traffic with such places and the minds that dwell there, you will not do so by reciting tidy cantrips. You will bargain. You will invite. You will bind yourself as surely as you bind them.

	\shadow{Rukhir clicks his beak and tilts one midnight eye at you. ``She thinks she is ready to be \emph{kept},'' he croaks, ``but does she know what it means to be \emph{kept}?''}

	These pages gather what the temple fires failed to burn: names that answer when called, diagrams that hum under the skin, rites that bear the weight of attention. Do not mistake them for common spellwork. A rite is a door you open with your own breath and blood, a vow you sign with the hours of your life.

	\section*{A Single Thread to Hold}
	You may walk with a companion from beyond, and you may keep a book of forbidden knowledge. \textbf{But you may only align with one Patron at a time.} The mind that tries to serve two masters will find itself torn and many patrons are quite jealous.

	\section*{The Price the Powers Prefer}
	Every calling across the veil is answered with a ledger. You may not hear the scratching of the quill at first, but the ink dries all the same. Each time you Invoke a rite, the one you call \emph{notices}. Their notice is not love and not hate; it is gravity.

	\begin{verse}
	Power borrowed is power tallied; \\
	A boon spent is a bell rung. \\
	When the bell is full of echoes, \\
	Something comes to listen.
	\end{verse}

	When your tallies swell, the world arranges a lesson. Perhaps the sea decides it is done being polite. Perhaps a truth hounds you down the alley you thought was yours. Do not panic; do not run. Kneel, count the echoes, pay what must be paid, and move with care.

	\whisper{Do not hide your costs from yourself. Hide them from your enemies, perhaps, but never from your own ledger.}

	\section*{On Patrons and Their Proper Address}
	Call them Witness or Geometry, Sea or Serpent, Fire or Law, Bone or Dream, the Elsewhere or the Red Court—names are handles, not cages. Treat each with the grammar they prefer. Order listens to proofs. Shadows favor vows made without torches. Seas respect the traveler who brings a tide of their own.

	\shadow{``Say a thing as if the world already agrees,'' Rukhir advises, ``and half the time the world is too proud to argue.''}

	Each rite you will meet below is written with \emph{The Weaving} (how to call), \emph{The Price} (what will be marked), and \emph{The Whisper} (what answer tends to come first). Read all three. Then read them again in a different mood.

	\section*{The Mercy of Severance}
There are rites to bind, and there is one to \emph{unbind}. I place it at the end of this gathering, under a black thread and a warning bell. We call it \emph{Cut the Knot of Patronage}. It is not a betrayal to sever a cord that is choking you. It is a kind of piety.

\shadow{``Mercy is a sharp thing,'' says Rukhir, preening. ``Only the kind learn to use it.''}

Do not attempt severance in mockery or haste. Bring witnesses who will speak your name without flinching. Bring an emblem of how you first said \emph{yes}. Bring enough quiet to hear the cord go slack.

\section*{How to Use What Follows}
Begin with one thread only. Learn how it pulls when you pull it. When you Invoke, mark; when you are marked, answer. If you carry a Codex, \emph{tend it}. If you walk with a Thiasos, \emph{guard it}. When attention gathers like storm-smell on stone, step under an eave and wait for the first fat drop. Then choose whether to dance, or to shut the door and count your candles.

\whisper{The spheres are not cruel. They are vast. Vastness feels like cruelty to those who will not look up. Look up. Then look back down and mind your footing.}

If you are still reading, you are ready enough to begin. Turn the page, seeker. The doors do not open themselves—but they do not stay closed for those who learn the grammar of keys.

\section{Design Premise}
Runekeepers leverage \textbf{Rites}---structured, named bargains with alien patrons. A Rite is not a spell; it does not use Channel/Weave. Instead, it runs on \textbf{Boons} and a \textbf{Patron Obligation Clock} that measures attention and debt.

\subsection*{Access Ladder for Rites}
\begin{description}
  \item[Low-Power Rites:] \textbf{Requires:} Either a Thiasos Bond \emph{or} a relevant Codex. \textbf{Invoke:} 1 Boon. \textbf{Obligation:} 4-segment clock.
  \item[Standard Rites:] \textbf{Requires:} Both a Thiasos Bond \emph{and} a relevant Codex. \textbf{Invoke:} 1 Boon. \textbf{Obligation:} 5–6 segments.
  \item[High-Power Rites:] \textbf{Requires:} Thiasos Bond, relevant Codex, plus \textbf{Tier III} standing with your Patron. \textbf{Invoke:} \textbf{2 Boons}. \textbf{Obligation:} 7–8 segments.
\end{description}

\paragraph{Single Patron Bond}
A character can only be bound to \textbf{one patron at a time}. Your Thiasos and Codex are intrinsically tied to this bond. To switch patrons, a character must undergo the \textbf{Rite of Severance} (\S\ref{subsec:severance}), a costly and dangerous process that severs the old bond before a new one can be formed.

\paragraph{Shared Limits}
A Rite's clock may gain at most \textbf{Tier +2 segments} per scene (Invoke + Push It + Backlash). If a single Rite marks \textbf{2+} segments in a scene, your connection strains visibly—your Thiasos shows distress or your Codex grows hot to the touch.

\paragraph{Stacking}
Bonuses from \emph{scene-long} Rites do not stack with other scene-long bonuses from the same Patron. Apply the single best effect.

\section{Core Procedure: Invoking a Rite}
\label{sec:rite-procedure}
\textbf{Invoke:} Spend the required \textbf{Boon(s)} and mark \textbf{+1 segment} on that Rite's Obligation Clock.

While the Rite is in effect this scene, you may mark at most \textbf{two more} segments from the following triggers (each may occur once/scene per Rite):
\begin{enumerate}
  \item \textbf{Push It:} You use the Rite's ``\emph{Push It}'' rider (the bolder, alternative clause). \emph{+1 segment}.
  \item \textbf{Backlash:} A roll \emph{aided by the Rite} results in a \emph{Miss} or \emph{Success \& Cost}. \emph{+1 segment}.
\end{enumerate}
\textbf{Cap:} A single Rite can add at most \textbf{Tier +2 segments} in a scene.

\textbf{Clearing Segments:}
\begin{itemize}
  \item During play, you may spend \textbf{1 Boon} to clear \textbf{1 segment} on a Rite.
  \item During Downtime, a \emph{Maintained} \textbf{Codex} clears \textbf{1 segment} from one associated Rite.
  \item Tending to your \textbf{Thiasos} during Downtime (a specific activity) can clear \textbf{1 segment} from one Rite.
  \item \textbf{Ritual of Appeasement:} Spend \textbf{1 Boon} and dedicate \textbf{Significant Time} in Downtime to clear \textbf{1 segment} from a Patron's Obligation Clock (once per Patron per session, requires 2+ segments).
\end{itemize}

\textbf{Attention:} When a Rite's Obligation Clock fills, the Patron's attention manifests as a narrative complication. Start a new clock for the "Patron's Lesson."

\section{Character Options}
\subsection{Talent: Thiasos Bond (Familiar)}
\label{subsec:thiasos-bond}
\textbf{Cost:} 6 XP \hfill \textbf{Type:} General Talent \\
\textbf{Effect:} You bind a \emph{Thiasos}---a supernatural familiar keyed to your Patron. It may Assist actions involving that Patron's Rites. The Thiasos follows Familiar rules for Exposure/Harm.\\
\textbf{Upkeep:} Requires occasional tending (roleplayed interaction) to remain receptive. A neglected Thiasos may refuse to Assist.\\
\textbf{Risk:} If a Rite marks 2+ segments in a scene, the Thiasos is endangered: mark \emph{Exposure +1} or suffer \emph{Harm 1} on it.

\subsection{Asset: Codex of the Binding Art}
\label{subsec:codex-asset}
\textbf{Type:} Asset (Minor 4 XP / Standard 8 XP) \\
\textbf{Scope:} A Codex is dedicated to a specific Patron. A Minor Codex covers 1-2 Rites; a Standard Codex covers 3-4 related Rites.\\
\textbf{Upkeep:} The Codex uses the Asset Condition track:
\begin{quote}
\textbf{Maintained} (full), \textbf{Neglected} (−1 die to Invokes), \textbf{Compromised} (unavailable).
\end{quote}
In Downtime, you may \emph{Study the Codex} to restore it to \emph{Maintained} and clear \textbf{1 segment} on an associated Rite.

\subsection{Rite Costs (Learning)}
\textbf{Learn a Rite:} 4–12 XP. Low-Power (4-5 XP), Standard (7-9 XP), High-Power (12-14 XP). Requires the appropriate access (Thiasos or Codex for Low; both for Standard; both + Tier III for High).

\section{Rites by Patron}
Each Patron's rites are organized into three tiers of power. All Invocations cost \textbf{1 Boon} unless noted. Push It effects now require a meaningful cost or limitation.

\input{patrons/index}


\section{Patron Relations and Jealousy}

Patrons do not share their chosen ones willingly. When a character bound to one Patron attempts to invoke rites of another, the first Patron's jealousy manifests as a complication (typically 1 SB) that reflects their domain and personality. This penalty does not apply to characters following the Invoker path, as they maintain no exclusive Patron relationships.

\subsection{Patron Relationship Dynamics}

\begin{table}[htbp]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Primary Patron} & \textbf{Relationship} & \textbf{Secondary Patron} & \textbf{Jealousy Response} \\
\hline
Any & Favor & Any Favored Patron & Never Jealous \\
\hline
Any & Loath & Any Loathed Patron & Always Jealous (+1 Obligation) \\
\hline
Any & Ambivalent & Any Ambivalent Patron & Sometimes Jealous (50\% chance, +1 Obligation) \\
\hline
\end{tabular}
\caption{Patron Jealousy Response Matrix}
\end{table}

\subsection{Key Patron Relationships}

\subsubsection{Major Conflicts}
\begin{itemize}
\item \textbf{Khemesh ↔ Raéyn}: Loath each other (depths vs. tides)
\item \textbf{Mor'iraath ↔ Mykkiel}: Loath each other (chaos vs. law)
\item \textbf{The Witness ↔ Maelstreus}: Loath (truth vs. deception)
\item \textbf{Ikasha ↔ Mykkiel}: Loath (shadow vs. law)
\item \textbf{Inaea ↔ Isoka}: Loath (rigid fate vs. fluid change)
\item \textbf{The Sealed Gate ↔ The Outsider}: Loath (containment vs. intrusion)
\item \textbf{The Carrion-King ↔ The Gallow's Bell}: Loath (natural decay vs. imposed judgment)
\item \textbf{Oath of Flame and Light ↔ Mor'iraath}: Loath (purity vs. corruption)
\item \textbf{Oath of Flame and Light ↔ The Outsider}: Loath (sacred order vs. chaos)
\end{itemize}

\subsubsection{Strong Alliances}
\begin{itemize}
\item \textbf{Ikasha → Raéyn}: Favor (shadow flows with tides)
\item \textbf{Isoka → Mor'iraath}: Favor (change feeds consumption)
\item \textbf{Inaea → Maelstreus}: Favor (webs connect to deals)
\item \textbf{The Traveler → Raéyn}: Favor (movement respects flux)
\item \textbf{Khemesh → The Carrion-King}: Favor (both embrace decay and depth)
\item \textbf{Varnek Karn → The Carrion-King}: Favor (both death-as-process)
\item \textbf{The Gallow's Bell → Mykkiel}: Favor (both concerned with justice/judgment)
\item \textbf{Oath of Flame and Light → Mykkiel}: Favor (both concerned with law and order)
\item \textbf{Oath of Flame and Light → The Gallow's Bell}: Favor (both concerned with justice and consequences)
\item \textbf{Oath of Flame and Light → The Sealed Gate}: Favor (both concerned with protection and sanctity)
\end{itemize}

\subsubsection{Complex Relationships}
\begin{itemize}
\item \textbf{Sacred Geometry ↔ Clockwork Demon}: Ambivalent (order vs. chaos, but both systematic)
\item \textbf{Varnek Karn ↔ The Carrion-King}: Ambivalent (both death, different aspects)
\item \textbf{Nidhoggr ↔ Valewood}: Ambivalent (both ancient, different memories)
\item \textbf{The Traveler ↔ The Ways Between}: Ambivalent (both concerned with paths, but one is a patron, one a realm)
\item \textbf{Mab ↔ The Outsider}: Ambivalent (both exist in spaces between, but Mab is court-structured while Outsider is chaotic)
\item \textbf{Sacred Geometry ↔ The Sealed Gate}: Ambivalent (both concerned with boundaries, but one creates order, one enforces barriers)
\item \textbf{Nidhoggr ↔ The Gallow's Bell}: Ambivalent (both concerned with the weight of the past, but one preserves ancient wisdom while other enforces consequences)
\item \textbf{Oath of Flame and Light ↔ Ikasha}: Ambivalent (light vs. shadow, but both concerned with truth - revealed vs. hidden)
\item \textbf{Oath of Flame and Light ↔ The Witness}: Ambivalent (both concerned with truth, but through different means - revelation vs. observation)
\end{itemize}

\subsection{Patron Jealousy Manifestations}

\subsubsection{Immediate Effects}
When jealousy is triggered:
\begin{itemize}
\item \textbf{+1 Obligation Segment} on current patron's clock
\item \textbf{Thiasos Distress}: Familiar shows agitation (-1 die assistance)
\item \textbf{Ritual Interference}: -1 die to rites involving offending patron
\item \textbf{GM Complication}: 1-2 SB of immediate narrative consequences
\end{itemize}

\subsubsection{Escalating Manifestations}
When Obligation Clock fills due to jealousy:

\paragraph{The Witness}
\begin{itemize}
\item Surveillance by unknown observers
\item Mysterious documents revealing duplicity
\item Whispers about secrets being shared
\end{itemize}

\paragraph{Sacred Geometry}
\begin{itemize}
\item Mathematical anomalies in rituals
\item Geometric patterns appearing accusingly
\item Structural instabilities in workspace
\end{itemize}

\paragraph{Ikasha}
\begin{itemize}
\item Shadows showing disapproval
\item Dreams of being watched
\item Whispers of betrayal in darkness
\end{itemize}

\paragraph{Khemesh}
\begin{itemize}
\item Oppressive weight sensations
\item Salt water appearing in wrong places
\item Deep rumbling sounds
\end{itemize}

\paragraph{Raéyn}
\begin{itemize}
\item Contrary winds/tides
\item Unpredictable weather
\item Messages containing warnings
\end{itemize}

\paragraph{Clockwork Demon}
\begin{itemize}
\item Device malfunctions during rituals
\item Gears grinding ominously
\item Steam hissing disapprovingly
\end{itemize}

\paragraph{Isoka}
\begin{itemize}
\item Unwanted physical transformations
\item Items changing unexpectedly
\item Paths shifting to complicate travel
\end{itemize}

\paragraph{Inaea}
\begin{itemize}
\item Coincidences with personal significance
\item Web patterns appearing accusingly
\item Strings tangling at crucial moments
\end{itemize}

\paragraph{Mor'iraath}
\begin{itemize}
\item Uncontrolled flames in rituals
\item Items heating uncomfortably
\item Smoke forming accusatory shapes
\end{itemize}

\paragraph{Mykkiel}
\begin{itemize}
\item Legal challenges to authority
\item Oaths requiring complex reaffirmation
\item Officials questioning credentials
\end{itemize}

\paragraph{Rite of the Borrowed Mantle (Advanced, 18 XP)} \emph{Duration: Instant; Range: Self; Stacking: No.}
\textbf{Materials:} A personal sacrifice worthy of the Patron (GM defined).
\textbf{Effect:} Before a roll, declare you are borrowing your Patron's mantle. You manifest a minor physical aspect. You may upgrade the effect level of your action by two steps (e.g., from Limited to Great) or add +3 dice to the roll.
\textbf{Push It:} The manifestation is so potent it affects the scene, creating a temporary environmental aspect related to your Patron that you can leverage once.
\emph{Requires: Familiar + Codex + Tier III + Special GM Approval \ (\textit{Invoke:} \textbf{3 Boons} and advance the Obligation Clock by 2 segments).}

\paragraph{Favor \& Attention}
\emph{Flavor.} A personified current of righteous light and purifying flame. It answers to oaths kept, mercy shown, and courage under shadow.  
\emph{Favor Suits.} \textbf{Hearts} (zeal, conscience), \textbf{Diamonds} (revelation, radiant signs), \textbf{Spades} (searing backlash).  
\emph{Patron's Symbol (examples).} A sun-sign stamped in cold iron; a glass ampoule holding a never-quenched ember; a burnished vow-ring engraved with a sunrise. You must possess the Symbol keyed to this Patron to Invoke its Rites.

\paragraph{Rite Tags}
Unless noted otherwise, each Rite follows the core procedure: \textbf{Invoke} (spend 1 Boon; mark +1 Obligation); once per scene the \textbf{Push It} rider may add +1 segment; a \textbf{Backlash} Miss or Success \& Cost on an aided roll adds +1 segment; max \textbf{Tier +3} segments/scene per Rite.

\paragraph{Access Ladder}
\emph{Low-Power} requires \textbf{Familiar + Patron's Symbol}. \;\;
\emph{Standard} requires \textbf{Familiar + Codex + Symbol}. \;\;
\emph{High-Power} requires \textbf{Familiar + Codex + Tier III + Symbol} and \textbf{Invoke 2 Boons}.

\bigskip

\section{Patron Manifestations}

When the Obligation Clock fills, a Patron will start to take interest in their charge. 

\begin{table}[htbp]
\centering
\begin{tabular}{>{\bfseries}p{0.8cm} p{4cm} p{8.5cm}}
\toprule
\textbf{d6} & \textbf{Manifestation Concept} & \textbf{Specific Patron Examples \& Lore Hooks} \\
\midrule
\textbf{1} & \textbf{The Subtle Shift} — Reality itself subtly realigns to reflect the Patron's nature. The change is environmental and persistent. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{The Witness:} Shadows hold unnaturally steady; mirrors show frozen flashes of the recent past. \\
        \textbf{The Sacred Geometry:} Angles drift toward pleasing ratios; spirals appear in wood grain and flowing water. \\
        \textbf{Oath of Flame and Light:} Candles within sight burn steady and bright; shadows retreat a finger's breadth from your skin; fresh wounds close into pale, clean scars overnight.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{2} & \textbf{The Unwelcome Gift} — The Patron bestows a power or insight that is useful but corrosive, demanding to be used. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{Ikasha:} You hear thoughts that hide in shadowed places; the whispers are accusatory and paranoid. \\
        \textbf{Mor'iraath:} Your touch smolders: you can kindle dry tinder at will, but delicate things you handle char and scar. \\
        \textbf{Oath of Flame and Light:} Your hands shed honest light: lies within arm's reach rasp and falter, yet small kindnesses you try to conceal blaze obvious to onlookers.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{3} & \textbf{The Hungry Messenger} — A creature or phenomenon aligned with the Patron dogs your steps, observing or making small demands. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{Varnek Karn:} An intelligent rat adopts you—bringing coins and trinkets, leaving neat arrangements of bones in your pack. \\
        \textbf{Mab:} A thorny vine sprouts where you bled; it blooms with weeping flowers that demand daily tending. \\
        \textbf{Khemesh:} Your footprints briefly fill with brackish water; gulls circle, and ropes creak where none are hung.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{4} & \textbf{The Reflected Flaw} — The Patron turns your own methods back on you, highlighting a hypocrisy or dependency. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{Mykkiel:} Casual promises bind literally; "a minute" becomes a geas until fulfilled. \\
        \textbf{Isoka:} A mask you wear becomes your public face; strangers insist the caricature is who you are. \\
        \textbf{Inaea:} A small deception you set in motion weeks ago snaps taut and entangles you in its fallout.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{5} & \textbf{The Echoing Demand} — The Patron requires a tangible task or sacrifice related to their domain. This is a direct quest. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{The Traveler:} Do not sleep under the same roof twice for a week, or dream of a road erasing your hometown from its maps. \\
        \textbf{Nidhoggr:} Recover and preserve a forgotten piece of history or suffer suffocating visions of being buried in time. \\
        \textbf{Maelstreus:} Broker a specific, morally gray bargain within a tenday or watch a profitable pact of yours unravel.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{6} & \textbf{The Personal Visitation} — The Patron, or a powerful emissary, appears in a dream or liminal space to deliver a warning or pivotal choice. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{The Outsider:} At the third hour before dawn, an extra, impossible corner opens in your room; a silent figure offers a key to a door that should not exist. \\
        \textbf{Oath of Flame and Light:} During a vigil, a figure of dawnfire offers to seal a vow: accept and gain strength to banish one named trespass, but relinquish a cherished grudge or secret. \\
        \textbf{Mab:} You attend a dream-feast where words are binding; break the table's rules and the consequences follow you into morning.
    \end{tabular} \\
\bottomrule
\end{tabular}
\end{table}

\section*{GM Guidance for Using the Table}

\begin{enumerate}
    \item \textbf{Weave, Don't Announce:} The manifestation should be introduced subtly. Don't say, ``The Witness is angry.'' Instead, describe how the PC notices that their shadow is moving a half-second too slow.

    \item \textbf{Start a Clock:} Each manifestation is the first segment of a new, typically 4-segment ``Patron's Attention'' clock. If this new clock fills, the consequences escalate from inconvenient to dangerous.

    \item \textbf{Offer a Remedy:} The manifestation should suggest a way to ``pay the debt'' or reduce the clock. For example, to appease \textbf{Varnek Karn}, the PC might need to give the rat a proper burial for a forgotten soul. Succeeding clears one segment.

    \item \textbf{Tie to the PC's Actions:} The chosen manifestation should feel like a direct consequence of how the PC has been using the Rites. A PC who used \textbf{Mor'iraath's} rites for destruction might get ``The Hungry Messenger'' (a small fire elemental that eats their belongings), while one who used them for passion might get ``The Reflected Flaw'' (their emotions become dangerously contagious).
\end{enumerate}

\section*{Imbunement Rites}

\begin{quote}\itshape
``He says the Gate wants the city sealed. I say it wants it opened---so that it may \emph{shut properly} when the time comes. The Gate smiles at us both.''
\end{quote}
\begin{flushright}
--- Scribe-Advocates of the Sealed Threshold, \textit{Third Petition}
\end{flushright}

\subsection*{Patron's Gift}
\textbf{Patron's Gift} (Free, Requires Thiasos )\\
Duration: Scene; Range: Touch; Stacking: No.\\
Materials: Personal item of significance.\\
Effect: Imbue one item with temporary magical power related to your Patron's domain. The item functions as a magical weapon(+1 Melee) and specialized tool (See Below) for this scene with +1 effect on relevant actions.\\
Push It: The item's power persists for one additional scene but marks +1 segment on your Obligation Clock.\\
Requires: Familiar (Invoke: 1 Boon).

\subsection*{Patron Imbunements}

\begin{table}[htbp]
\centering
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{>{\bfseries}p{3.2cm} >{\raggedright\arraybackslash}p{6.1cm} >{\raggedright\arraybackslash}p{6.1cm}}
\toprule
\textbf{Patron} & \textbf{Basic Effect} & \textbf{Push It Effect} \\
\midrule
The Witness & Item bears a truth-mark — \textbf{+1 die} to \emph{Investigation/Notice} & Item reveals one hidden truth about the current target or environment \\
The Sacred Geometry & Item settles into perfect ratio — \textbf{+1 die} to precision, timing, or measured tasks & Item ignores the first penalty from terrain or unstable footing on one action \\
Ikasha & Item takes on umbral aspect — \textbf{+1 die} to \emph{Stealth/Deceive} & Item (or its wielder's hand) may pass through one solid barrier unimpeded (once) \\
Khemesh & Item gains crushing undertow — \textbf{+1 effect} on raw force or break actions & On one hit, item ignores armor/hardness \\
Raéyn & Item flows with the turn — \textbf{+1 die} to adaptation, navigation, or weathered travel & Treat one environmental hazard or shift as one step less severe (once) \\
The Clockwork Demon & Item ticks with proscribed precision — \textbf{+1 die} to \emph{Craft/Repair} & Item functions despite one missing or broken component (once) \\
Isoka & Item loosens bonds — \textbf{+1 die} to \emph{Escape/Transform} actions & Shed one minor physical or mental condition (once) \\
Inaea & Item threads the social weave — \textbf{+1 die} to \emph{Influence/Manipulate} & Expose one relevant connection, obligation, or leverage between NPCs \\
Mor'iraath & Item wreathed in hungry flame — \textbf{+1 effect} on destructive actions & Flammables ignite on contact with a successful hit (this scene) \\
Mykkiel & Item bears the seal — \textbf{+1 die} to \emph{Authority/Command} & Compel minor, immediate compliance from a lawful subject (once) \\
Varnek Karn & Item tastes of the barrow — \textbf{+1 effect} vs. living targets & Ignore 1 \emph{Fatigue} on a single exertion (once) \\
Mab & Item is fae-touched — \textbf{+1 die} to \emph{Charm/Deceive} & Conjure a minor illusion or misdirection (once) \\
Nidhoggr & Item hums with antiquity — \textbf{+1 die} vs. ancient or supernatural wards & Reveal one pertinent historical fact or true name about the target \\
The Outsider & Item phases at the edge — \textbf{+1 die} vs. material barriers or confinement & Briefly exist "between" to bypass one obstacle or occupy two points for an instant (once) \\
Maelstreus & Item tempts the ledger — \textbf{+1 effect} on gambles, bargains, or risky leverage & Convert one \emph{Miss} into \emph{Success \& Cost} (once) \\
The Traveler & Item keeps the way — \textbf{+1 die} to travel, navigation, or pursuit/escape & Treat one difficult-terrain or engagement penalty as null for a single movement \\
The Carrion-King & Item hastens decay — \textbf{+1 effect} on sabotage or corrosion & Cause one organic object to rot, split, or foul beyond use (once) \\
Oath of Flame and Light & Item bears dawnfire — \textbf{+1 die} to \emph{defend/heal} or when striking undead, demons, or outsiders & On a successful hit against such a foe, \emph{Turn} it (it withdraws, DV=Tier) \emph{or} it immediately generates 2 SB (\emph{Spades}) \\
\bottomrule
\end{tabular}
\end{table}

\section{Ritual of Severance: Exorcism of Patronage}
\label{subsec:exorcism}
\textbf{Name:} \emph{Cut the Knot of Patronage} \quad \textbf{Type:} Ritual (Extended) \\
\textbf{Who:} The warlock, \emph{plus} 1--3 assistants (any Arts; distinct procedures). \\
\textbf{Materials:} A mirror veiled in salt-thread, anointed Codex page or Thiasos token, a personal sacrifice tied to the Patron's domain. \\
\textbf{Procedure:} Over an extended scene, the circle denies \emph{names, signs, and vectors} that bind the Patron. Each helper offers a distinct symbolic inversion tied to their skill.
\begin{itemize}
  \item \textbf{Cost:} The primary invoker spends \textbf{2 Boons}. Mark \emph{Exposure +1}. Each helper risks 1 Fatigue on a \emph{Miss}.
  \item \textbf{Test:} GM sets DV by entanglement (typ. DV 2–3). On success, \textbf{clear the Rite's Obligation Clock} to 0 and \textbf{choose one}: (a) \emph{Dismiss} your Thiasos \underline{or} (b) \emph{Seal} your Codex (cannot be used for that Patron until re-consecrated in Downtime).
  \item \textbf{On Success \& Cost:} As above, but the GM immediately starts a fading \emph{Echo} thread keyed to the Patron (2–3 SB over the next scene).
  \item \textbf{On Miss:} The knot tightens. Mark \emph{+2 segments} on the active Patron clock, and the GM introduces a \emph{Geas}-like demand or intrusion.
\end{itemize}
\textbf{Aftermath:} You may switch paths (Codex $\leftrightarrow$ Thiasos) once the severance is complete and the inert resource is relinquished or reconsecrated in Downtime.


\section*{Invoker Path: Ritual Access}

Instead of the Familiar + Codex approach, an Invoker accesses Rites through \textbf{Symbol + known Rite}. All Invoker Rites are \textbf{rituals} and take \textbf{Significant time} (typically 10–30 minutes; GM sets fiction). Invoker Rites \emph{do not} use Push It. The caster must openly display the relevant \textbf{Patron's Symbol} throughout.

\paragraph{Attention \& Obligation (Invoker)}
Completing an Invoker ritual \textbf{marks +1 segment} on that Rite's \emph{Obligation Clock} (even if outside a scene). This replaces any Boon cost. If you \emph{Crack the Seal} (see below) to act instantly, mark more.

\subsubsection*{Ritual Novice (4 XP)}
\textbf{Effect:} You may perform any \textbf{Low-Power} Rite you \emph{know} as a sacred ritual. The Rite's duration becomes \textbf{Extended} unless it is already longer.\\
\textbf{Limitations:} No Push It; suitable environment; visible Symbol throughout.\\
\textbf{Crack the Seal (Instant Cast):} Immediately resolve the Rite without time. \textbf{Set the Symbol to \emph{Compromised}} and \textbf{mark +2 segments} on that Rite's Obligation. The GM may immediately spend \textbf{1 SB} related to the Patron's domain.\\
\textbf{Jealousy:} Invoking a Patron you are not pledged to \textbf{also} marks \textbf{+1 segment} on that Patron's \emph{Obligation}.

\subsubsection*{Ritual Adept (6 XP)} 
\textbf{Effect:} As above, and you may perform \textbf{Standard} Rites you \emph{know}.\\
\textbf{Requires:} Ritual Novice.

\subsubsection*{Ritual Master (8 XP)} 
\textbf{Effect:} As above, and you may perform \textbf{High-Power} Rites you \emph{know}.\\
\textbf{Instant Cast (High-Power):} When you \emph{Crack the Seal} on a High-Power Rite, \textbf{mark +3 segments} on that Rite's Obligation and set the Symbol to \emph{Compromised}.\\
\textbf{Requires:} Ritual Adept.

\paragraph{Concurrent Bindings (cap)}
You may sustain at most \textbf{Spirit} ongoing Invoker rituals at once. Beginning a new one forces the oldest to lapse or immediately marks \textbf{+1 segment} on its Obligation (your choice).

\paragraph*{Symbol Vulnerability }

When a Patron's Symbol is compromised through damage, theft, or magical interference, it cannot be used for Rite invocation until repaired. Compromised Symbols:
\begin{itemize}
\item Cannot be used for 24 hours
\item Require Downtime repair (1 day) or XP (2 XP) to restore
\item Risk of permanent damage on 3+ compromises
\end{itemize}

The GM may introduce complications when a Symbol is compromised, such as attracting unwanted attention from entities aligned with that Patron or causing minor magical malfunctions in the bearer's other equipment.

\paragraph{Multi-Symbol Risk}
You may carry multiple Symbols (one per Patron). If you carry \textbf{3+} different Patrons' Symbols, the first Invoker ritual you complete each scene marks \textbf{+1 additional segment} on its Obligation (the choir gets loud).

\subsection*{Patron's Symbol (Minor Asset)}\index{Patron's Symbol}\index{Assets}

\textbf{Type:} Minor Asset (4 XP) \hfill \textbf{Scope:} One specific Patron

\textbf{Requirement.} To Invoke any Rite of a Patron via the Invoker path, you must \textbf{possess that Patron's Symbol}. You may carry only \textbf{one} Symbol per Patron (but may hold Symbols of different Patrons).

\textbf{What it is.} A consecrated token keyed to a Patron's office (sign, seal, relic). It anchors names and permissions.

\textbf{Condition \& Upkeep.}
\begin{itemize}
  \item \textbf{Maintained:} Functions normally.
  \item \textbf{Neglected:} Ritual completes but \textbf{mark +1 extra Obligation segment} on that Rite.
  \item \textbf{Compromised:} Inert—cannot be used until \textbf{restored}.
\end{itemize}

\textbf{Crack the Seal (Instant Cast).} As part of an Invoker Rite, you may \textbf{immediately} resolve the effect by \textbf{setting the Symbol to \emph{Compromised}} and marking \textbf{+2 segments} (\textbf{+3} if High-Power) on that Rite's Obligation. The GM may spend \textbf{1 SB} on-theme. \emph{This does not remove the Asset.}

\textbf{Restore a Symbol (Downtime).} Repair a \emph{Compromised} Symbol with a Downtime action and a fitting test (DV 3 or by fiction). On success, set to \emph{Maintained}; on a shaky result, it returns as \emph{Neglected}. Alternatively, spend \textbf{1 XP} to fully restore it between sessions.

\textbf{On-Screen Activation (once/scene).} Presenting/working the Symbol costs \textbf{1 Boon}: \textbf{+1 die} \emph{or} \textbf{improve Position by one step} for negotiation, warding, or parley with that Patron's agents/influences.

\textbf{Off-Screen (1/session).} Petition a sign, omen, or minor audience relevant to the Patron (GM sets limits).

\paragraph{Loss, Theft, and Forgeries}
A stolen/forged Symbol cannot Invoke your Rites, but can misdirect omens. A false Sign treated as real begins a short "False Sign" complication thread.

\begin{table}[htbp]
\centering
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{>{\bfseries}p{4.0cm} >{\raggedright\arraybackslash}p{10.8cm}}
\toprule
\textbf{Patron} & \textbf{Symbol Examples (choose or craft one)} \\
\midrule
The Witness & Obsidian eye-sigil pendant scored with a single lash; a silvered mirror shard wrapped in black wire; a wax seal-stamp bearing an open eye, carried on a cord. \\
The Sacred Geometry & Brass heptagram rule or compass with seven-point tooth; bone tablet inlaid with a golden-ratio spiral; plumb-bob knotted with a proof written in red thread. \\
Ikasha & Vial of soot-oil that drinks light; knot of black silk whose shadow forms a sigil; fingerbone ring lacquered matte, always cool to the touch. \\
Khemesh & Salt-etched iron chain link; lead sounder-weight engraved with abyssal curls; barnacle-bitten coin drilled and worn as a token. \\
Raéyn & Sea-glass disk notched for tides; storm-twined rope knot crusted with salt; glass ampoule of rainwater gathered at three crossings. \\
The Clockwork Demon & Gear tooth sealed in a tube of red oil; mainspring coil etched with forbidden numerals; rivet of unknown alloy stamped with a demon's stamp. \\
Isoka & Shedding of snakeskin braided into a wrist loop; quicksilver ampoule that gleams like a new skin; brass torque engraved with interlaced serpents. \\
Inaea & Spool of spider-silk thread around a bone bobbin; silver ring of interlocking knots; amber bead with a tiny inclusion "caught" at center. \\
Mor'iraath & Coal set in an iron claw; glass ampoule holding a live ember that never cools; charred vow-strip bound with copper wire. \\
Mykkiel & Cold-iron seal matrix engraved with a balance; square rule stamped with a code number; parchment writ-tag tied with red cord and wax. \\
Varnek Karn (The Carrion-King) & Ossuary bead rosary threaded with grave-dust; carved phalanx-bone tally; coin of bone and obsidian fused and drilled. \\
Mab & Hawthorn thorn wrapped in silver wire; mirror shard backed with green felt; acorn cup lined with red silk, worn as a charm. \\
Nidhoggr & Fossil tooth shard on a black thong; river-stone that has never seen sunlight, kept in a dark pouch; obsidian spindle with a hairline flaw. \\
The Outsider & Cracked mirror shard laced with silver wire; key with no teeth and no ward; crystal prism that throws an extra, impossible ray. \\
Maelstreus & Signet ring bearing infernal script; miniature contract scroll sealed in red glass; coin scored in halves and linked by a fine chain. \\
The Traveler & Road-nail wrapped in waythread; waystone pebble tied with ribbon and dusted in road-salt; brass compass missing its needle that nevertheless points. \\
Oath of Flame and Light & Cold-iron sun-stamp; glass ampoule with a consecrated spark; vow-ring engraved with a sunrise and the bearer's true name. \\
\bottomrule
\end{tabular}
\end{table}
\subsection*{Rite Access Ladders}

\paragraph{Invoker Path}
\begin{itemize}
  \item \textbf{Novice:} Low-Power Rites you know as rituals (\emph{+1 segment on completion})
  \item \textbf{Adept:} Standard Rites you know as rituals (\emph{+1 segment})
  \item \textbf{Master:} High-Power Rites you know as rituals (\emph{+1 segment})
  \item \textbf{Instant Cast (any tier):} \emph{Crack the Seal} (\emph{Symbol $\rightarrow$ Compromised}; \emph{+2 segments}, or \emph{+3} if High-Power; GM may spend 1 SB)
\end{itemize}

\subsubsection{Invoker Path Limitations}

The Invoker path trades immediacy and flexibility for reliability and accessibility. These limitations ensure that Invokers remain balanced against traditional Rite users while maintaining their unique identity as ritual specialists.

\paragraph{Casting Time Constraints} All Invoker Rites are \textbf{rituals} requiring \textbf{Significant Time} (typically 10-30 minutes). The GM sets the specific fiction requirements based on the Rite's nature and power level. Emergency casting is possible through the \textbf{Crack the Seal} mechanic, but at significant cost.

\paragraph{No Push It Benefits} Invoker Rites \textbf{cannot utilize Push It effects}. The ritualistic nature of Invoker magic requires precise, predetermined procedures that do not allow for the improvisational enhancements available to bonded Rite users.

\paragraph{Environmental Requirements} Rituals require a \textbf{suitable environment} appropriate to the Rite's nature. Complex rituals may require specific ceremonial spaces, ritual purity, or freedom from interruption. The GM determines environmental suitability based on fiction.

\paragraph{Symbol Display Requirement} The \textbf{Patron's Symbol must be openly displayed} throughout the ritual performance. Concealing or hiding the Symbol disrupts the ritual connection and prevents the Rite from functioning.

\paragraph{Concurrent Ritual Limit} An Invoker may sustain at most \textbf{Spirit} ongoing Invoker rituals simultaneously. Beginning a new ritual forces the oldest active ritual to lapse or immediately marks \textbf{+1 segment} on its Obligation Clock (caster's choice).

\paragraph{Multi-Symbol Overload} Carrying \textbf{4 or more different Patrons' Symbols} creates "choir" effects - the first Invoker ritual completed each scene marks \textbf{+1 additional segment} on its Obligation Clock as conflicting divine harmonics interfere with precise ritual work.

\paragraph{Attention & Obligation Generation} Completing an Invoker ritual marks \textbf{+1 segment} on that Rite's Obligation Clock \textbf{even outside scene context}. This represents the constant attention required to maintain ritual connections with powerful entities through mere symbols rather than bonded relationships.

\paragraph{Crack the Seal Consequences} Using the instant cast mechanism immediately sets the Symbol to \textbf{Compromised} and marks \textbf{+2 segments} (+3 for High-Power) on the Rite's Obligation Clock. The GM may immediately spend \textbf{1 SB} related to the Patron's domain theme, representing the instability of rushed ritual work.

\chapter{Closing Words}

\section*{The Price of Knowledge}

\lettrine[lines=3]{A}{}s I set down my pen and watch the ink dry on these final pages, I am struck by the weight of what I have recorded. Each ritual, each warning, each whispered secret represents not just power gained, but power that has been tested, measured, and found wanting. For knowledge without wisdom is merely dangerous, and power without purpose is merely destruction.

\shadow{Rukhir has been silent for hours now, his black eyes reflecting the candlelight like twin stars. Even he, ancient as he is, seems wearied by the cataloging of so much darkness.}

In my years of walking the shadow-paths, I have learned that the greatest magic is not in the spells we cast or the rituals we perform, but in the choices we make when the candles burn low and the whispers grow loud. Every ritual in this grimoire has been tested—not just for its efficacy, but for its cost. And make no mistake, dear reader, each one demands a price that would make merchants weep and kings tremble.

The Bonding Ceremony can unite hearts, but it can also chain souls. The Sanctuary Consecration can offer protection, but it can also become a prison. The Memory Reconstruction can reveal truth, but it can also unleash madness. Each path we walk in shadow leads both toward power and toward peril, and the wise witch knows that the difference lies not in avoiding the darkness, but in understanding it.

\section*{A Word to the Keeper}

If you have read this far, then perhaps you are ready for what lies ahead. But readiness is not the same as wisdom, and wisdom is not the same as safety. The rituals and \emph{Rites} I have shared with you are not toys for idle amusement or tools for petty revenge. They are conversations with powers who remember their due. Each Rite is a door you open with a \textbf{Symbol} or a carefully kept \textbf{Codex}, and every door opens \emph{both} ways.

\shadow{"Knowledge is a burden," Rukhir finally speaks, his voice like rustling leaves. "And every burden leaves marks upon the bearer."}

Use these workings with respect for those who taught me, for those who suffered for this knowledge, and for those who may yet suffer because of it. Shadow-magic is not evil in itself—nor are the Patrons simple saints or monsters. They are \emph{offices} of the world: Witness and Law, Tide and Road, Ember and Depth, the silk of Fate and the hiss of Change. In the hands of the unworthy, any Rite becomes a weapon; in the hands of the desperate, even a blessing can burn.

Remember always that Ikasha, She Who Sleeps Within Shadow, is not a goddess of chaos or evil. She is the patron of those who understand that truth often hides in darkness, that power comes at a cost, and that sometimes the only way to fight the monsters is to borrow their grammar. Others answer, too: the Geometry to proofs, the Traveler to thresholds, the Oath-bound Light to vows kept. They will \emph{listen}—and they will \emph{remember}.

\section*{The Path Forward}

For those who would continue down the shadow-roads, I offer these final pieces of advice:

First, never trust completely. Even the most faithful familiar has its own agenda, and even the most courteous Patron is jealous of attention. Trust, but verify. Love, but protect. Hope, but prepare. When you \emph{Invoke}, state your price and hear theirs.

Second, remember that power is a responsibility. The ability to influence others, to bend position, to peer into hidden places—these are not gifts to be used lightly. With each \emph{Rite} you perform, with each seal you break, you write your name deeper into the ledger of \emph{Obligation}. Keep your \textbf{Symbol} tended, your \textbf{Codex} clean, and your promises cleaner still.

Third, know when to walk away. There are secrets that should remain buried, powers that should remain unfound, and roads that should remain unopened. When Law and Tide argue, choose your shore. When the Witness offers truth that will unmake a friend, choose which world you will live in afterward. Wisdom is as much in knowing what not to do as in knowing what to do.

\shadow{Rukhir spreads his wings, and in the candlelight, they seem to hold fragments of every shadow I have ever seen. "The end of one path," he says, "is always the beginning of another."}

If the cords you have tied begin to choke you, there is a knife for knots: the severance rite. Use it without shame. Consent is a law deeper than any Patron's pride.

Finally, remember that you are not alone. The Tulkani have a saying: "No shadow falls without light to cast it." In every darkness, there is the potential for illumination. In every ending, a seed of beginning. The Rites in this grimoire are not the end of your journey—they are merely the next step upon a path that has no final destination, a correspondence you keep with the world's older names.

\section*{The Eternal Whisper}

As you close this book and prepare for whatever comes next, remember the fundamental truth Ikasha taught me in my first communion with her shadow:

\begin{verse}
\textit{The deepest shadows} \\
\textit{Are cast by the brightest flames.} \\
\textit{The truest vows} \\
\textit{Are written in signs and scars.} \\
\textit{The strongest Rites} \\
\textit{Are paid for twice—} \\
\textit{Once in the asking,} \\
\textit{And again when the Patron answers.}
\end{verse}

May your shadows be deep, your flames be bright, and your wisdom be greater than your power. Tend your Symbols. Mind your ledgers. Choose your Patrons as carefully as you choose your friends.

\shadow{Rukhir caws once, sharp and final, and I know that our work here is done.}

Until we meet again in the spaces between the words, dear reader, may the shadows keep you safe, and may the light never burn too brightly for your eyes.

\begin{center}
\textit{Finis}
\end{center}

\shadow{But in the world of Rites and patrons, nothing is ever truly finished. Only... paused.}

\end{document}
