\section{The Lethai Ghe'hai — Elite Operatives of the Silk Courts}

\subsection{Domain and Epithets}

The Lethai do not send armies. They send individuals. The Ghe'hai—a term that resists direct translation, meaning roughly "those who walk the thread between" —are elite operatives trained from adolescence in the arts of combat, infiltration, diplomacy, and elimination. They are not assassins in the common sense; they are problem-solvers who happen to be lethal. A Ghe'hai can kill a target in a crowded ballroom and walk out without anyone knowing, or they can talk a warlord into surrendering without drawing steel. Their Patron is not a single entity but a syncretic weave—drawing from Inaea (webs of obligation), Isoka (transformation of identity), and Ikasha (shadow and patience), with a thread of the Witness (truth as testimony).

\subsection{Cultural Origins}

The Ghe'hai tradition emerged from the Crimson Valley and the Lethai-thora cities:

\begin{itemize}
\item In the \textbf{Crimson Valley}, the Ghe'hai serve as neutral arbiters, sent to resolve disputes that cannot be settled by law. Their word is binding. Their silence is a verdict.
\item In \textbf{Lethai-thora} enclaves, the Ghe'hai are the Hidden Hand, known to exist but never seen. Their symbols are a silk mask, a silver needle, and a single strand of red thread.
\item Among the \textbf{Lethai-ar}, the Ghe'hai are the Unseen Court, invoked only when all other options have failed. They do not speak of their missions. Neither does anyone else.
\end{itemize}

\subsection{Warrior Archetype: The Silk-Blade}

A Runic Warrior of the Lethai syncretic tradition is not a knight or a berserker. They are a \textit{Silk-Blade}—an operative who moves through society as easily as through shadow, who can kill with a whisper or negotiate with a smile, who wears masks as easily as others wear clothes. Their Codex is a silk mask or a cord of red thread. Their Thiasos is a spider-spirit that weaves threads between people, a memory-echo that records confessions, or a sentient blade wrapped in silk that leaves no mark.

\subsection{Codex Form}

The Silk-Blade's Codex is a testament to disguise and discretion:

\begin{itemize}
\item \textbf{Silk Mask:} A mask of woven silk that shifts patterns. It can assume the appearance of any face the Runekeeper has touched. The mask does not sweat. It does not age.
\item \textbf{Red Thread Cord:} A braided cord of crimson silk, tied around the wrist. Each knot is a mission completed. The cord tightens when the Runekeeper is recognized.
\item \textbf{Silver Needle:} A needle of silver, cold to the touch. It can stitch wounds—or stitch lips shut. It never dulls.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Spider-Spirit:} A translucent spider the size of a hand. It weaves threads between people—obligation, memory, debt. It can sense hidden relationships within Near range.
\item \textbf{Memory-Echo:} A spirit that cannot speak, but can replay sounds it has witnessed—a confession, a password, a name spoken in the dark. It records everything. It forgets nothing.
\item \textbf{The Silk Mask (Sentient):} A mask that speaks in a neutral, androgynous voice. It can assume the appearance of any face the Runekeeper has touched (once per session, for one scene). The mask remembers every face it has worn.
\end{itemize}

\subsection{Core Mechanic: The Unseen Witness}

When the Silk-Blade is not recognized (disguised, hidden, or simply forgettable), they gain +1 die to all social and stealth rolls. The witness does not see.

When the Silk-Blade eliminates a target without being identified, they gain 1 Boon. This Boon can only be used on rolls involving escape, disguise, or misdirection.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Velvet Touch (4 XP)}

\textbf{Tags:} [STEALTH], [DECEPTION], [SHADOW]

\textbf{Threshold:} A silk ribbon touched to lips.

Your movements become silent. Your presence becomes forgettable. Gain +1 die to a single Stealth or Subterfuge roll. If you succeed, witnesses suffer -2 dice to recall your description.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Unseen Question (5 XP)}

\textbf{Tags:} [COMMAND], [FEAR], [TRUTH]

\textbf{Threshold:} A circle drawn with silk thread while focusing on a specific person.

Ask one question of a target within the zone. They may either answer truthfully (the GM must give a true answer, though it may be cryptic) or mark 1 Fatigue and gain a Guilty Tell Condition.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Mask of Many Faces (8 XP)}

\textbf{Tags:} [TRANSFORM], [VEIL], [DECEPTION]

\textbf{Threshold:} A silk mask and a personal token from the person whose face you wish to wear.

Assume the appearance of any humanoid you have touched for one scene. Your voice, build, and scent shift to match. You gain +2 dice to Deception rolls to pass as that person. Those who knew the original well may test Resolve (DV 4) to notice something is wrong.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Silk Binding (7 XP)}

\textbf{Tags:} [BIND], [TRAP], [RESTRAINT]

\textbf{Threshold:} A length of silk cord, woven with intent.

Lay an invisible snare—a tripwire, a binding thread, a closing loop. The first creature to trigger it must test Body + Athletics (DV 4) or be immobilised for one exchange. The snare lasts for the scene or until triggered.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Unspoken Question (12 XP)}

\textbf{Tags:} [MIND], [TRUTH], [FEAR]

\textbf{Threshold:} A reflective surface (mirror, still water, polished metal) and a drop of the target's blood or a strand of their hair.

The target tests Resolve (DV 4) or suffers -2 dice to rolls involving memory, deception, or resisting interrogation for the scene. If they fail, you may ask one specific, factual question about something they know, and they must answer truthfully or suffer 1 SB (Hearts) as the memory is forcibly drawn forth.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Ghe'hai's Escape (6 XP)}

\textbf{Requirement:} Lethai syncretic tradition

Once per session, when you are cornered or capture is inevitable, you may vanish, reappearing in a safe location of the GM's choosing. You leave behind a token—a silk mask, a red thread, a single word—that can be used to track you. The GM gains 2 SB.

\subsubsection{The Unseen Blade (4 XP)}

\textbf{Requirement:} Lethai syncretic tradition

When you attack from a hidden position or from a crowd, your target suffers -1 die to defend. They do not see the threat until it is too late. If you are disguised, the target cannot describe you afterward.

\subsubsection{Silk Memory (8 XP)}

\textbf{Requirement:} Lethai syncretic tradition, Witnessed Scar

Once per session, you may touch a silk cord or mask and recall a perfect sensory memory of any scene you have witnessed while wearing it—sounds, smells, temperatures, the exact words spoken. The memory may be used as testimony (in a court that respects Lethai witness) or as a clue.

\subsection{Roleplaying a Silk-Blade}

\begin{itemize}
\item You do not have a single face. You have many. You are not sure which one is real.
\item You may be called to infiltrate, to observe, to eliminate, or to negotiate. The mission is the mask. The mask is the mission.
\item You may struggle with identity. When you have worn too many faces, you may forget which one is yours. The silk does not judge. It simply covers.
\item You may be asked to stop—to keep one face, to keep one name, to be one person. The Ghe'hai does not retire. The Ghe'hai is reassigned.
\end{itemize}

\subsection{Sample NPC: The Mask of the Crimson Vigil}

\textbf{Archetype:} Silk-Blade

\textbf{Patron:} Lethai syncretic

\textbf{Codex:} A silk mask that shifts patterns

\textbf{Thiasos:} A spider-spirit that weaves threads between people

\textbf{Personality:} Calm, precise, unknowable. They have worn nine faces this year. They are not sure which one is theirs.

\textbf{Conflict:} A mission requires them to eliminate someone they once loved—under a different face, a different name, a different life. The mask remembers. The heart is not sure.

\subsection{GM Guidance: Intrusions for the Ghe'hai}

\begin{itemize}
\item A former mission resurfaces. Someone has recognized the Silk-Blade—not their face, but their gait, their silence, their way of holding a cup.
\item The Silk-Blade's mask begins to itch. The face underneath is not their own. They cannot remove the mask.
\item A thread the Silk-Blade tied between two people tightens unexpectedly. A debt is coming due. The debtor does not know who holds the cord.
\item The Silk-Blade dreams of a room with nine masks on the wall. One is missing. The missing mask is their own face.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
