% -\textit{- coding: utf-8 -}-
% Chapter: Character Advancement
% Comprehensive Resource Guide for Fate's Edge

\chapter{Character Advancement}\label{chap:character_advancement}
\index{Advancement}\index{Experience Points}\index{Character Creation}

\section{Introduction}\label{sec:advancement_intro}
\index{Advancement!Overview}

Character advancement in \textbf{Fate's Edge} reflects meaningful growth in capability and standing. XP is awarded for table-facing accomplishments, hard choices, and the dramatic friction that defines your story. Advancement should always reinforce narrative identity: what drives, obligations, and affinities are shaping the character?\index{XP!Awards}\index{Narrative identity}

\section{Starting Character Builds}\label{sec:starting_builds}
\index{Character Creation!Starting builds}

\subsection{Standard Starting XP}\label{subsec:standard_starting}
\index{Character Creation!32 XP}

Players begin with \textbf{32 XP} to allocate during initial character creation. This represents a balanced baseline for competent starting characters.

\subsection{Enhanced Starting Options}\label{subsec:enhanced_starting}
\index{Character Creation!Enhanced builds}\index{Bonds}\index{Complications}

Players may exceed the standard 32 XP build through narrative engagement:
\begin{itemize}
    \item \textbf{Bonds:} Up to two player-defined mutual bonds may be taken for +2 XP total (+1 XP each).\index{Bonds!Starting bonus}
    \item \textbf{Complications:} Up to two initial complications may be accepted for +4 XP total (+2 XP each).\index{Complications!Starting bonus}
\end{itemize}
This allows a maximum starting build of \textbf{36XP}. Players are encouraged to aim for 32 XP and use bonds/complications to enhance characterization rather than pure mechanical optimisation.\index{Character Creation!36XP cap}

\subsection{Initial Complications Reminder}\label{subsec:initial_complications}
\index{Complications!Initial}

For each initial complication taken at character creation:
\begin{itemize}
    \item Start each scene with +1 banked SB per character with initial complications.\index{Story Beats!Banked}
    \item These complications remain until cleared through play or character development.\index{Complications!Clearing}
\end{itemize}

\subsection{Small Folk of the Threshold}\label{subsec:small_folk_threshold}
\index{Aelaerem}\index{Aelinnel}\index{Small Folk}

The Aelaerem and Aelinnel are diminutive peoples attuned to liminal spaces and hidden ways. Their stature grants them agility and subtlety, though at the cost of bearing heavy arms or armor.

\begin{itemize}
    \item \textbf{Restriction:} Cannot use \emph{Heavy Armor} or \emph{Heavy Weapons}.
    \item \textbf{Bonus:} Gain +1 \emph{Position} when Dodging or Resisting Knockback, and +1 die on \emph{Hide} or \emph{Evasion} rolls made while in cover.
\end{itemize}

Their presence in the world is often underestimated, but their knack for slipping unseen through thresholds and enduring where others falter has earned them a quiet reverence.

\section{Player Archetypes}\label{sec:player_archetypes}
\index{Archetypes}\index{Build styles}

Not every group plays the same way. These archetypes describe how players typically spend XP and seek spotlight.

\subsection{The Solo}\label{subsec:solo_archetype}
\index{Archetypes!Solo}

\begin{description}
    \item[\textbf{Definition}] Invests XP primarily in Attributes and Skills. Minimal followers, minimal holdings. All power is on the character sheet.\index{Attributes}\index{Skills}
    \item[\textbf{Typical XP Spread}] 70--90\% Self; 0--10\% On-screen help; 0--20\% Off-screen assets.\index{XP!Spreads}
    \item[\textbf{Strengths}] Consistent scene impact; few moving parts; resilient to follower loss.\index{Followers!Loss resilience}
    \item[\textbf{Risks}] Limited fiction reach between sessions; can stall when problems demand logistics or networks.\index{Logistics}\index{Networks}
    \item[\textbf{Starting Focus}] Attributes 2--3, Skills 1--2, minimal assets.\index{Assets!Minimal}
\end{description}

\subsection{The Mixed Player}\label{subsec:mixed_archetype}
\index{Archetypes!Mixed}

\begin{description}
    \item[\textbf{Definition}] Splits XP between self-growth and one or two meaningful assets (a small follower or a reliable holding).\index{Assets}\index{Followers}
    \item[\textbf{Typical XP Spread}] 50--65\% Self; 15--25\% On-screen help; 15--25\% Off-screen assets.\index{XP!Spreads}
    \item[\textbf{Strengths}] Versatile: credible in scenes and has a lever for special problems.\index{Versatility}
    \item[\textbf{Risks}] Upkeep pressure; helper can be targeted when the GM spends 2+ Story Beats.\index{Upkeep}\index{Story Beats!GM spends}
    \item[\textbf{Starting Focus}] Balanced approach with one minor asset or low-cap follower.\index{Assets!Minor}\index{Followers!Cap}
\end{description}

\subsection{The Mastermind}\label{subsec:mastermind_archetype}
\index{Archetypes!Mastermind}

\begin{description}
    \item[\textbf{Definition}] Prioritises followers / cadres / familiars and off-screen networks. The character sheet is the hub of a larger apparatus.\index{Networks}\index{Cadres}\index{Familiars}
    \item[\textbf{Typical XP Spread}] 25--40\% Self; 35--55\% On-screen help; 20--40\% Off-screen assets.\index{XP!Spreads}
    \item[\textbf{Strengths}] Scene control via assistance; strategic reach between sessions; strong heist / social-planning play.\index{Teamwork}\index{Planning}
    \item[\textbf{Risks}] Dependency on assist lanes; followers can be endangered on 2+ SB spends; upkeep pressure.\index{Assists}\index{Followers!Risk}
    \item[\textbf{Starting Focus}] Significant investment in followers or major assets from the beginning.\index{Assets!Major}
\end{description}

\section{XP Costs and Advancement}\label{sec:xp_costs}
\index{XP!Costs}\index{Advancement!Costs}

\subsection{Core Character Improvements}\label{subsec:core_improvements}
\index{Attributes!Costs}\index{Skills!Costs}\index{Followers!Costs}

\begin{longtable}{p{3.5cm} p{9cm}}
\caption{Core Improvement Costs}\label{tab:core_costs}\\
\toprule
\textbf{Improvement} & \textbf{Cost and Downtime} \\
\midrule
\endfirsthead
\multicolumn{2}{c}%
{{\bfseries \tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Improvement} & \textbf{Cost and Downtime} \\
\midrule
\endhead
\bottomrule
\endfoot
Attribute increase & Cost = new rating \(\times 3\) XP; downtime = new rating (days). \\
Skill increase & Cost = new level \(\times 2\) XP; downtime = new level (days). \\
On-screen Follower & Cost = Cap\(^{2}\) XP; 1--3 days to recruit and brief. \\
\end{longtable}

\subsection{Off-Screen Assets}\label{subsec:off_screen_assets_costs}
\index{Assets!Costs}

\begin{longtable}{p{3.5cm} p{9cm}}
\caption{Asset Costs}\label{tab:asset_costs}\\
\toprule
\textbf{Asset Type} & \textbf{Cost and Downtime} \\
\midrule
Minor Asset & 4 XP; 1 day. \\
Standard Asset & 8 XP; 1 week. \\
Major Asset & 12 XP; 1 month. \\
Artifact Asset & 16+ XP; campaign effort. \\
\bottomrule
\end{longtable}

\subsection{Talent Costs}\label{subsec:talent_costs}
\index{Talents!Costs}

\begin{longtable}{p{3.5cm} p{9cm}}
\caption{Talent Costs}\label{tab:talent_costs}\\
\toprule
\textbf{Talent Tier} & \textbf{Cost} \\
\midrule
Minor Talents & 2--3 XP (small situational bonus). \\
Major Talents & 4--6 XP (strong upgrade, permanent effect in a niche). \\
Prestige Talents & 7--10 XP (campaign-defining, breaks fundamental limits). \\
Epic Talents & 11+ XP (legendary, story-shaping). \\
\bottomrule
\end{longtable}

\section{Tiers of Reputation}\label{sec:reputation_tiers}
\index{Reputation Tiers}

Though there are no character levels, total XP spent creates soft ``tiers'' that shape how the world responds to characters.\index{XP!Totals}

\begin{longtable}{p{3.5cm} p{9cm}}
\caption{Reputation Tiers}\label{tab:reputation_tiers}\\
\toprule
\textbf{Tier} & \textbf{Effect} \\
\midrule
Tier I -- Novice (0--40 XP) & Local reputation; prestige abilities locked. \\
Tier II -- Seasoned (41--90 XP) & Regional notice; prestige abilities may unlock. \\
Tier III -- Veteran (91--150 XP) & National influence; second follower slot suggested. \\
Tier IV -- Paragon (151--220 XP) & Movers and shakers; rivals emerge to challenge. \\
Tier V -- Mythic (221+ XP) & Legendary status; kingdoms and cults respond directly. \\
\bottomrule
\end{longtable}

\section{Prestige Abilities and Talents (Examples)}\label{sec:prestige_abilities}
\index{Talents}\index{Prestige Abilities}

Prestige abilities are narrative milestones unlocked through mastery or story events. They are priced at 6+ XP and include campaign-shaping effects.\index{Milestones}

\subsection{General Talents}\label{subsec:general_talents}
\index{Talents!General}

\begin{description}
    \item[\textbf{Battle Instincts}] (6 XP): Once per scene, re-roll a failed defense roll.\index{Battle Instincts}
    \item[\textbf{Silver Tongue}] (4 XP): Gain +1 die when persuading or deceiving through speech.\index{Silver Tongue}
    \item[\textbf{Iron Stomach}] (6 XP): Immune to mundane poisons and spoiled food; halve complications from toxic sources.\index{Iron Stomach}
    \item[\textbf{Exceptional Coordination}] (8 XP): One follower can provide +4 assist dice (exception to the +3 cap).\index{Exceptional Coordination}\index{Assists!Maximum}
\end{description}

\paragraph{Subtle Casting (Major Talent -- 8 XP)}
\textit{Prerequisite:} Lore 3+, Performance 2+ \textbf{or} Runekeeper with Codex.

\noindent \textbf{Effect:} Make a \textbf{Performance + Lore} roll to quietly cast a spell, invoke a Rite, or sing a Cantos against DV (Tier). If successful, the casting does \textit{not} generate Story Beats on the Channel or initial roll. Any SB generated are \textbf{banked by the GM} and applied at dramatically appropriate moments.

\noindent \textbf{Limitations:}
\begin{itemize}
    \item Cannot be used for \textbf{Great} or \textbf{Extreme} Tier effects.
    \item The \emph{Weave} phase (if applicable) still generates normal SB.
\end{itemize}

\paragraph{Backstab (Major Talent -- 8 XP)}
\textit{Prerequisite:} Stealth 2+, Melee 2+, Light weapon.

\noindent \textbf{Effect:} When you attack an \textbf{Unaware} or \textbf{Engaged} foe from \textbf{Stealth}, deal \emph{+1 Harm} and ignore \emph{1 point of their Armor}.

\noindent \textbf{Definitions:}
\begin{itemize}
    \item \textbf{Unaware:} Target not aware of your presence or hostile intent.
    \item \textbf{Engaged:} Target is taking an \emph{Attack} action against another character, or casting a spell that targets another character.
\end{itemize}
\noindent \textbf{Limit:} Once per scene. On a miss, you are \emph{Exposed} -- drop to Desperate Position or mark 1 Harm.

\paragraph{Shadow Dance (Synergy Talent -- 10 XP)}
\textit{Prerequisite:} Backstab, Stealth 3+, Mobility 2+.

\noindent \textbf{Effect:} After a successful \textbf{Backstab}, test \textbf{Stealth} vs. DV (Tier). On success, re-enter Stealth and either clear 1 Fatigue or improve Position +1. On failure, remain Exposed and mark 1 Fatigue. May only chain once per scene.

\paragraph{Deathblow (Capstone Talent -- 12 XP)}
\textit{Prerequisite:} Shadow Dance, Stealth 4+, Melee/Ranged 3+.

\noindent \textbf{Effect:} Strike from Dominant Position or after re-entering Stealth, declare a Deathblow. On a hit, deal \emph{double harm}. If the attack incapacitates the target, attempt a free Stealth test to vanish. On a miss, become Exposed (Desperate Position and mark 1 Harm). Once per scene; may mark 1 Fatigue to attempt a second time.

\subsection{Cultural and Ancestral Talents}\label{subsec:cultural_talents}
\index{Talents!Cultural}

\subsubsection{Humans}\label{subsubsec:human_talents}
\index{Humans!Talents}

\begin{description}
    \item[\textbf{Versatile}] (3 XP): Once per session, swap one Skill die for another you lack.\index{Versatile (Human)}
    \item[\textbf{Guild Ties}] (7 XP; Presence 2+): Call in favors from a guild or order once per session.\index{Guild Ties (Human)}\index{Presence}
    \item[\textbf{Banner-Bearer}] (15 XP; Presence 4+, Resolve 3): Inspire an allied unit; once per battle, erase 2 complications for the group.\index{Banner-Bearer (Human)}\index{Resolve}
\end{description}

\subsubsection{Dwarves (Aeler)}\label{subsubsec:dwarf_talents}
\index{Dwarves!Talents}\index{Aeler}

\begin{description}
    \item[\textbf{Stone-Sense}] (5 XP): Detect flaws in stone; +1 die to underground \emph{Lore} or \emph{Survival} checks involving stonework.\index{Stone-Sense (Dwarf)}\index{Lore}\index{Survival}
    \item[\textbf{Spirit Shield}] (8 XP; Spirit 3+): Commune with ancestors; once per session, block 1 complication for an ally.\index{Spirit Shield (Dwarf)}\index{Spirit}
    \item[\textbf{Forge-Patriarch}] (18 XP; Body 4+, Lore 3): Found a forge-citadel; gain loyal smiths and engineers (Cap 5 followers).\index{Forge-Patriarch (Dwarf)}\index{Followers!Cap 5}\index{Body}
\end{description}

\subsubsection{Wood Elves}\label{subsubsec:wood_elf_talents}
\index{Wood Elves!Talents}

\begin{description}
    \item[\textbf{Backlash Soothing}] (5 XP): Once per session, cancel 1 Backlash die in natural terrain.\index{Backlash Soothing (Wood Elf)}\index{Backlash}
    \item[\textbf{Ranger's Step}] (9 XP; Wits 3+, Stealth 2): Move unseen in forests; treat terrain penalties as one step lower.\index{Ranger's Step (Wood Elf)}\index{Wits}\index{Stealth}
    \item[\textbf{Wild Speaker}] (18 XP; Spirit 5+, Survival 3): Speak to beasts and trees; once per session, command local wildlife as allies.\index{Wild Speaker (Wood Elf)}\index{Survival}
\end{description}

\subsubsection{High Elves}\label{subsubsec:high_elf_talents}
\index{High Elves!Talents}

\begin{description}
    \item[\textbf{Lorekeeper}] (4 XP): Recall obscure history or magic without rolling once per session.\index{Lorekeeper (High Elf)}
    \item[\textbf{Weave Anchor}] (8 XP; Wits 3+, Arcana 3): Reduce magical Backlash by 1 when casting Intricate spells.\index{Weave Anchor (High Elf)}\index{Arcana}
    \item[\textbf{Ways-Walker's Step}] (20 XP; Wits 5, Arcana 4): Once per arc, turn a complication into a boon; GM banks 2 SB.\index{Ways-Walker's Step (High Elf)}\index{Omens}\index{Difficulty Value}
\end{description}

\subsubsection{Ykrul}\label{subsubsec:ykrul_talents}
\index{Ykrul!Talents}

\begin{description}
    \item[\textbf{Blood Frenzy}] (4 XP; Body 2+): When reduced to half health, gain +1 die on melee rolls.\index{Blood Frenzy (Ykrul)}\index{Body}
    \item[\textbf{Blood Memory}] (7 XP; Body 3+): After battle, gain 1 temporary Skill die reflecting a foe's tactics in the next scene.\index{Blood Memory (Ykrul)}\index{Skills}
    \item[\textbf{Warglord}] (18 XP; Body 5, Sway 3): Once per campaign, unify scattered warbands; start Logistics and Grudge timers.\index{Warglord (Ykrul)}\index{Sway}\index{Timers}
\end{description}

\section{Starting Build Examples}\label{sec:starting_examples}
\index{Character Creation!Examples}\index{Build examples}

\subsection{Solo Build (30--32 XP)}\label{subsec:solo_example}
\index{Archetypes!Solo!Example}

\begin{itemize}
    \item Body 3 (9 XP), Wits 2 (6 XP), Spirit 2 (6 XP).
    \item Skills: Combat 2 (4 XP), Stealth 1 (2 XP), Survival 1 (2 XP).
    \item Minor Talent: Iron Stomach (3 XP).
    \item \textbf{Total:} 32 XP (requires 1 bond or minor complication).
\end{itemize}

\subsection{Mixed Build (32 XP)}\label{subsec:mixed_example}
\index{Archetypes!Mixed!Example}

\begin{itemize}
    \item Attributes: Body 2 (6 XP), Wits 2 (6 XP), Spirit 2 (6 XP).
    \item Skills: Sway 2 (4 XP), Lore 1 (2 XP).
    \item Minor Asset: Safehouse (4 XP).
    \item Cap 2 Follower (4 XP).
    \item \textbf{Total:} 32 XP.
\end{itemize}

\subsection{Mastermind Build (36XP)}\label{subsec:mastermind_example}
\index{Archetypes!Mastermind!Example}

\begin{itemize}
    \item Attributes: Body 2 (6 XP), Wits 3 (9 XP), Spirit 2 (6 XP).
    \item Skills: Sway 2 (4 XP).
    \item Cap 3 Follower (9 XP).
    \item \textbf{Total:} 36XP (requires 2 bonds or complications).
\end{itemize}

\section{Advancement Guidance}\label{sec:advancement_guidance}
\index{Advancement!Guidance}

\begin{itemize}
    \item Advancement should always reinforce narrative identity and character goals.\index{Narrative identity}
    \item Consider how new abilities fit the character's story and relationships.\index{Relationships}
    \item Balance immediate power with long-term narrative potential.\index{Power curve}
    \item Work with the GM to ensure advancements make sense in the campaign context.\index{GM guidance}
\end{itemize}

\section{Optional Rule: Multi-Character Followers}\label{sec:multi_character_followers}
\index{Followers!Multi-character}\index{Optional rules}

\subsection{Overview}
A player may control multiple characters by linking them as mutual Followers. This option also covers absent players`` characters when consent is given.

\subsection{Core Mechanics}
\begin{itemize}
    \item \textbf{Cap Tier +1:} A Follower may advance up to one Tier higher than their leader's current Tier.
    \item \textbf{Multi-Character Play:} A player may designate two PCs as Followers of each other. Only one is the \emph{active} PC at a time; the other functions as a Follower.
    \item \textbf{Absent PCs:} With consent, an absent player's PC may be run as a temporary Follower at Cap Tier +1.
\end{itemize}

\subsection{Switching Active Characters}\label{subsec:switching_active}
\index{Followers!Switching}

\begin{itemize}
    \item \textbf{Switch Cost:} When switching, the previously active PC immediately becomes a Follower of the new active PC at Cap Tier +1.
    \item \textbf{Condition Gate:} A PC cannot be switched \emph{to} if Compromised, Seized, or otherwise narratively unfit for active duty.
    \item \textbf{Narrative Justification:} The fiction must support the switch (returning from a journey, recovering from harm, taking point for a specific task).
    \item \textbf{Strategic Choice:} Switching trades autonomy for support---gaining a powerful Follower but losing a full PC.
\end{itemize}

\subsection{Balance Notes}
\begin{itemize}
    \item Prevents abuse by making switching itself the cost.
    \item Maintains narrative logic -- leadership shifts naturally reshape party structure.
    \item Creates strategic depth -- switching is a meaningful tactical decision.
    \item Keeps all characters relevant, encouraging upkeep even when not active.
\end{itemize}

% End of chapter