% -\textit{- coding: utf-8 -}-
% Chapter: GM Reference Tables
% Comprehensive Resource Guide for Fate's Edge

\chapter{GM Reference Tables}\label{chap:gm_reference}
\index{GM reference}\index{Tables}

\definecolor{FeAccent}{HTML}{3F6E8C}
\definecolor{FeRow}{HTML}{F5F7FA}
\newcommand{\seg}{\(\square\)}
\newcommand{\segf}{\(\blacksquare\)}
\newcommand{\segt}{\(\boxtimes\)}
\renewcommand{\arraystretch}{1.15}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcommand{\feTableStart}{\rowcolors{2}{FeRow}{white}\small}
\newcommand{\feTableEnd}{\rowcolors{2}{}{}}

\section*{Core Resolution Cycle}\label{sec:core_cycle_ref}
\index{Core resolution!GM reference}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l Y@{}}
\toprule
\textbf{Core Resolution (at a glance)} & Declare \textbf{Action \& Approach} $\rightarrow$ Set \textbf{DV} $\rightarrow$ \textbf{Roll d10 pool} $\rightarrow$ Count \textbf{Successes (6+)} and \textbf{1s = SB} $\rightarrow$ Apply \textbf{Position} reroll rule $\rightarrow$ Compare to \textbf{DV} and resolve. \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\section*{Outcome Matrix}\label{sec:outcome_matrix_ref}
\index{Outcome Matrix}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l l Y@{}}
\toprule
\textbf{Result} & \textbf{Boons} & \textbf{SB (GM)} & \textbf{GM Guidance} \\
\midrule
$S \geq DV$, $C=0$ & 0 & Spend 1s as needed & \textbf{Clean Success}: grant intent crisply. \\
$S \geq DV$, $C>0$ & 0 & Spend/bank 1s & \textbf{Success \& Cost}: intent + immediate complication. \\
$0<S<DV$ & 1 & Spend/bank 1s & \textbf{Partial}: proportional progress or open a fork/complication. \\
$S=0$ & 2 & Spend/bank 1s & \textbf{Miss}: no progress; escalate timers/pressure. \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\section*{Position, DV, Effect, Criticals}\label{sec:position_ref}
\index{Position}\index{Difficulty Value (DV)}\index{Effect}\index{Critical success}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l l Y@{}}
\toprule
\textbf{Position} & \textbf{Mechanical Edge} & \textbf{Consequence Severity} & \textbf{Typical} \\
\midrule
Dominant  & Re-roll \emph{one failure} (1--5) & Minor & Ambush, superior prep, leverage. \\
Controlled& (No built-in reroll)               & Moderate & Even footing; ready. \\
Desperate & Must re-roll \emph{one success} (6--10) & Severe & Cornered, outmatched, wounded. \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}>{\centering\arraybackslash}p{1.2cm} l Y@{}}
\toprule
\textbf{DV} & \textbf{Difficulty} & \textbf{Use Cases} \\
\midrule
2 & Routine   & Clear intent, modest stakes, controlled environment. \\
3 & Pressured & Time limits, mild resistance, partial information. \\
4 & Hard      & Hostile conditions, active opposition, precision. \\
5+& Extreme   & Stacked constraints, dramatic failure risk. \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l Y@{}}
\toprule
\textbf{Effect} & \textbf{Impact} & \textbf{Examples} \\
\midrule
Limited  & Minor     & Scratch damage; inch forward; narrow opening. \\
Standard & Expected  & Normal damage; hit the objective as framed. \\
Great    & Major     & Significant harm; rapid progress; surplus benefit. \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l Y@{}}
\toprule
\textbf{Criticals \& Counting} & \textbf{10s count as \emph{two} successes each. 1s always create SB (even on rerolls). Criticals \emph{do not} auto-succeed a DV if $S<DV$. Apply Position rerolls \emph{after} the initial count; re-rolled dice can add SB if they show 1.} \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\section*{Story Beats (SB) and Boons}\label{sec:sb_boons_ref}
\index{Story Beats}\index{Boons}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}>{\centering\arraybackslash}p{1.6cm} l Y@{}}
\toprule
\textbf{SB Cost} & \textbf{Scale} & \textbf{Examples} \\
\midrule
1 SB  & Minor pressure   & Noise/trace; +time; chip Position; strain a resource. \\
2 SB  & Moderate setback & Alarm; lose cover; lesser foe arrives; close a route. \\
3 SB  & Serious trouble  & Reinforcements; key gear breaks; front twists. \\
4+ SB & Major turn       & Authority arrives; scene shift; disaster trigger. \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l Y@{}}
\toprule
\textbf{Boon Cost} & \textbf{Effect} & \textbf{Limits} \\
\midrule
1 Boon & Re-roll one die (after Position) & Once per action. \\
1 Boon & Improve Position by 1 step       & Max +1/action. \\
1 Boon & Activate on-screen Asset         & Plausibility test if needed. \\
2 Boons& Convert to 1 XP                  & Max 2 XP/session via conversion. \\
Var.   & Power talents/rites              & As specified. \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\section*{Timers, Harm, Fatigue}\label{sec:timers_harm_ref}
\index{Timers}\index{Harm}\index{Fatigue}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l >{\centering\arraybackslash}p{1.2cm} *{6}{>{\centering\arraybackslash}p{0.8cm}} Y@{}}
\toprule
\textbf{Clock} & \textbf{Size} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{Trigger/Effect} \\
\midrule
\_\_\_\_ & 4/6/8 & \seg & \seg & \seg & \seg & \seg & \seg & What advances it / What it does when filled. \\
\_\_\_\_ & 4/6/8 & \seg & \seg & \seg & \seg & \seg & \seg & \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\begin{tcolorbox}[title={Fatigue \& Harm Reference},colback=FeRow,colframe=FeAccent,breakable]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Fatigue Level} & \textbf{Effect} \\
\midrule
0 & Fresh: no effect. \\
1 & Winded: downgrade one Controlled roll to Desperate per scene. \\
2 & Strained: Controlled rolls add +1 SB on 1s. \\
3 & Exhausted: downgrade one Controlled roll to Desperate per scene. \\
4 (Max) & Collapse: further exertion = DV 3 Body test or Severe Harm. \\
\bottomrule
\end{tabularx}

\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Harm Level} & \textbf{Casting Impact} \\
\midrule
Minor & Concentration taxed; channeled effects DV +1. \\
Moderate & Risk of disruption: roll DV 2 to maintain channels. \\
Severe & Channel breaks; freeform casting auto-generates +1 SB. \\
Critical & No further casting; Patron may impose terms for reprieve. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\section*{Difficulty Value (DV) by Tier}\label{sec:dv_by_tier}
\index{Difficulty Value (DV)!by Tier}

\begin{center}
\feTableStart
\begin{longtable}{c c p{8cm}}
\caption{DV by Character Tier}\label{tab:dv_tier}\\
\toprule
\textbf{Tier} & \textbf{Base DV} & \textbf{Example Challenge} \\
\midrule
\endfirsthead
\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Tier} & \textbf{Base DV} & \textbf{Example Challenge} \\
\midrule
\endhead
\bottomrule
\endfoot
I  & 3 & Local threat / novice test. \\
II & 4 & Veteran foe or skilled task. \\
III& 5 & Elite / magical challenge. \\
IV & 6 & Mythic or cosmic threat. \\
V  & 7+ & Legendary, world-shifting. \\
\end{longtable}
\end{center}
\noindent \textbf{Position Modifiers:} Dominant --1, Controlled ±0, Desperate +1.

\section*{Common Action DVs (Quick Reference)}\label{sec:common_dvs}
\index{Difficulty Value (DV)!common actions}

\begin{center}
\feTableStart
\begin{longtable}{l l l}
\toprule
\textbf{Action Type} & \textbf{Typical DV} & \textbf{Notes} \\
\midrule
Basic Attack (melee/ranged) & 2 & Even footing. \\
Maneuver (change Position, create advantage) & 2--3 & Set up an ally. \\
Active Defense (parry, block, evade) & 3 & Under pressure. \\
Complex Action (multi-target, strong control) & 4 & Area attack, powerful spell. \\
High-Risk Gambit & 5+ & Desperate, extreme precision. \\
\bottomrule
\end{longtable}
\end{center}

\section*{Magic Backlash (Condensed)}\label{sec:backlash_ref}
\index{Backlash!quick reference}

\begin{center}
\feTableStart
\begin{longtable}{>{\bfseries}l l l}
\toprule
\textbf{Element} & \textbf{Minor Backlash} & \textbf{Major Backlash} \\
\midrule
Earth / Fate   & Position --1 or \textsc{Encumbered} & Timer +1/2 (Collapse) or \textsc{Pinned} \\
Fire / Life    & Effect --1 or \textsc{Singed} & Timer +1 (Fire) or 1 Harm \\
Air / Luck     & Position --1 or Timer +1/2 & Immediate Cost (lose tool) or +1 SB \\
Water / Obishaal & Effect --1 or \textsc{Waterlogged} & Timer +1/2 (Flood) or intrusion \\
\bottomrule
\end{longtable}
\end{center}

\section*{Obligation and Patrons}\label{sec:obligation_ref}
\index{Obligation!quick reference}

\begin{center}
\feTableStart
\begin{longtable}{l l}
\toprule
\textbf{Action} & \textbf{Obligation Cost} \\
\midrule
Invoke Low Rite (Runekeeper) & +1 \\
Push It (amplify Rite) & +1 additional \\
Invoke Standard Rite & +2 \\
Crack the Seal (Instant Invoker) & +2 (+3 for High Rites) \\
Patron's Gift (Imbuement) & 0 (with Thiasos) \\
Borrowed Grace (Invoker talent) & +1 \\
\bottomrule
\end{longtable}
\end{center}
\noindent Obligation Capacity = Spirit + Presence. Exceeding capacity = +1 Fatigue per segment over.

\section*{Equipment Quick Reference}\label{sec:equipment_ref}
\index{Equipment!quick reference}

\subsection*{Weapon Modifiers (Melee)}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l c c l@{}}
\toprule
\textbf{Weight} & \textbf{Close} & \textbf{Near} & \textbf{Notes} \\
\midrule
Light  & +2d & +1d & Fast, concealable. \\
Medium & +1d & +2d & Set 1/scene or --1d first attack. \\
Heavy  & --1d & +3d & Set 1/scene or --2d first attack. \\
\bottomrule
\end{tabularx}
\feTableEnd

\subsection*{Ranged Tempo}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l c c c l@{}}
\toprule
\textbf{Weight} & \textbf{Tempo} & \textbf{Close} & \textbf{Near} & \textbf{Far} \\
\midrule
Light  & Fast      & Controlled & +1d & --- \\
Medium & Standard  & Desperate  & +2d & +1d \\
Heavy  & Slow      & Desperate  & +1d & +3d \\
\bottomrule
\end{tabularx}
\meTableEnd

\subsection*{Armor Conversion}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l c c l@{}}
\toprule
\textbf{Armor} & \textbf{Conversion} & \textbf{Penalty} & \textbf{Notes} \\
\midrule
Light  & 1 Harm $\rightarrow$ 1 Fatigue & --- & \\
Medium & 2 Harm $\rightarrow$ 1 Fatigue & --1d physical & \\
Heavy  & 3 Harm $\rightarrow$ 2 Fatigue & --2d physical, no sprint & \\
\bottomrule
\end{tabularx}
\feTableEnd

\section*{Travel Timer Sizes}\label{sec:travel_timers}
\index{Travel!timer sizes}

\begin{center}
\feTableStart
\begin{longtable}{l c p{8cm}}
\toprule
\textbf{Card Rank} & \textbf{Timer Size} & \textbf{Typical Journey} \\
\midrule
2--5   & 4 segments & Short, straightforward. \\
6--10  & 6 segments & Standard journey. \\
J, Q, K & 8 segments & Extended or complex. \\
Ace   & 10 segments & Epic or highly dangerous. \\
\bottomrule
\end{longtable}
\end{center}

\section*{Skill List (Abridged)}\label{sec:skills_ref}
\index{Skills!list}

\feTableStart
\begin{tabularx}{\linewidth}{@{}l Y@{}}
\toprule
\textbf{Skill} & \textbf{Typical Uses} \\
\midrule
Arcana & Magic theory, wards, ritual preparation. \\
Athletics & Running, climbing, swimming, jumping. \\
Brawl & Unarmed combat, grappling, improvised fighting. \\
Command & Leadership, intimidation, issuing orders. \\
Craft & Making, mending, construction. \\
Deception & Lying, bluffing, misdirection. \\
Insight & Reading emotions, detecting lies, motives. \\
Investigation & Research, deduction, analyzing clues. \\
Lore & History, culture, religion, ancient sites. \\
Medicine & Healing, first aid, diagnosis. \\
Melee & Close combat with weapons. \\
Notice & Perception, spotting ambushes, hidden details. \\
Performance & Acting, music, oration. \\
Ranged & Bows, thrown, firearms. \\
Stealth & Moving unseen, hiding. \\
Subterfuge & Casing, impersonation, forgery, sleight of hand. \\
Survival & Wilderness navigation, foraging, tracking. \\
Sway & Persuasion, negotiation, charm. \\
Tinker & Lockpicking, traps, mechanical devices. \\
\bottomrule
\end{tabularx}
\feTableEnd

\section*{Talent Cost Tiers}\label{sec:talent_costs_ref}
\index{Talents!cost tiers}

\begin{center}
\feTableStart
\begin{longtable}{l l p{8cm}}
\toprule
\textbf{Tier} & \textbf{Cost} & \textbf{Scope} \\
\midrule
Minor   & 2--3 XP & Small situational bonus, once per scene. \\
Major   & 4--6 XP & Strong upgrade, permanent effect in a niche. \\
Prestige& 7--10 XP & Campaign-defining, breaks fundamental limits. \\
Epic    & 11+ XP & Legendary, story-shaping. \\
\bottomrule
\end{longtable}
\end{center}

\section*{Asset and Follower Costs}\label{sec:asset_costs_ref}
\index{Assets!costs}\index{Followers!costs}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l Y@{}}
\toprule
\textbf{Asset Type} & \textbf{XP Cost} & \textbf{Establishment Time} \\
\midrule
Minor  & 4 XP & 1 day \\
Standard & 8 XP & 1 week \\
Major  & 12 XP & 1 month \\
\bottomrule
\end{tabularx}

\begin{tabularx}{\linewidth}{@{}l l Y@{}}
\toprule
\textbf{Follower Cap} & \textbf{XP Cost} & \textbf{Assist Dice} \\
\midrule
1 & 1 XP & +1d (if relevant skill) \\
2 & 4 XP & +2d \\
3 & 9 XP & +3d \\
4 & 16 XP & +3d (cap applies) \\
5 & 25 XP & +3d (cap applies) \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\section*{XP Awards per Session}\label{sec:xp_awards_ref}
\index{Experience Points!awards}

\begin{center}
\feTableStart
\begin{longtable}{l l}
\toprule
\textbf{Award Type} & \textbf{XP Range} \\
\midrule
Table Attendance & +2 XP \\
Major Objective & +2--4 XP \\
Discovery & +1--2 XP \\
Hard Choice & +1--2 XP \\
Complication Spotlight & +1--3 XP \\
Bond-Driven Play & +1--2 XP \\
GM Curveball & +0--3 XP \\
Arc Completion (major) & +8--12 XP \\
\bottomrule
\end{longtable}
\end{center}

\section*{Reputation Tiers}\label{sec:reputation_tiers_ref}
\index{Reputation Tiers}

\begin{center}
\feTableStart
\begin{longtable}{l l l}
\toprule
\textbf{Tier} & \textbf{XP Range} & \textbf{Influence} \\
\midrule
I -- Novice   & 0--40 XP   & Local reputation. \\
II -- Seasoned & 41--90 XP & Regional notice. \\
III -- Veteran & 91--150 XP & National influence. \\
IV -- Paragon & 151--220 XP & Continental reputation. \\
V -- Mythic   & 221+ XP   & Legendary, world-changing. \\
\bottomrule
\end{longtable}
\end{center}

% End of chapter