\chapter{The Black River at Dawn}

\section*{Roran did not sleep the night after the trial.}

He sat on the cold stone bench of the Queen's garrison, the silver gauntlet heavy on his left hand, the locket warm against his palm on the right.

The gauntlet still smelled of the forge where it had been shaped—hammer strikes, the hiss of cooled steel, the ledger-books that recorded every blade, every horse, every ounce of grain as if they were the only things that gave a life meaning. The locket was soft, its surface worn smooth by years of being pressed against a mother's chest.

He had not opened it. He could not.

Whatever lay inside—a portrait, a lock of hair, a written name—would be the first proof that his mother had existed outside the cold arithmetic of his father's ledgers. Outside the columns of debts and dues.

He was afraid.

Not of the locket's contents. Of what it would cost him to keep it.

\vspace{1em}

At the third bell, the garrison's watch changed. The corridors fell quiet. He rose, the gauntlet clinking faintly against the stone.

He moved through the sleeping city. His boots were silent on cobblestones that had once echoed with the march of soldiers. He passed the Silent Stair, where the wind whistled through broken stones like a lament. He passed the chapel where Alain the Cantor, voice raw from years of singing the \textit{Planctus Armatus}, dreamed of feathers and forgotten songs.

The gates stood open. The war was over. The peace was young. The watch, trained for vigilance, had learned to sleep again.

The Black River lay ahead—a slow, dark ribbon drinking the first grey light of dawn.

\vspace{1em}

He knelt at the bank. Cold water seeped through his breeches. The air smelled of iron and wet stone, of rain that had fallen days ago and would fall again.

He lifted the locket to his chest. Felt his own heartbeat against its cool metal.

He opened it.

\vspace{1em}

Inside was not a portrait.

It was a braid. A single, thin braid of auburn hair, tied with a scrap of red thread. The same thread Bryn had once looped between a chair leg and a pot handle. The same thread Margot had used to knit a blanket for a child she would never see grow.

The braid was warm.

He pressed it to his lips. Closed his eyes. Listened.

Silence.

And then—

\begin{center}
\textit{``Aldric.''}
\end{center}

The voice came from the water. Or from the mist. Or from the hollow of his own chest. It was not his mother's voice alone.

He looked up.

She stood on the far bank.

Young. Younger than him. Younger than Eleanor. The age she had been when the birthing fever took her. Her hair fell loose, auburn and unbound. Her face held no sorrow—only a quiet certainty. A red blanket draped over her shoulders like a cloak.

\begin{center}
\textit{``Mother,'' he said.}
\end{center}

She smiled.

\begin{quote}
\textit{``You have carried the weight long enough. The locket was never meant to be a burden. It was meant to be a bridge.''}
\end{quote}

She stepped toward the water's edge. The river did not part. It simply accepted her. She stood on the surface as if the current had always been solid beneath her feet.

\begin{quote}
\textit{``Lay it at the bank. Tie it with the thread. Let the water carry it where it needs to go.''}
\end{quote}

\begin{quote}
\textit{``Where?''}
\end{quote}

\begin{quote}
\textit{``To her. To Bryn. She kept it for forty years. She never asked for thanks. But she deserves to know it reached you.''}
\end{quote}

He felt the old name strike him like a blow.

\begin{quote}
\textit{``That name forsook me. I am Roran now. That is the name I chose when I left the ledger behind.''}
\end{quote}

Her eyes softened. Not with disappointment. With understanding.

\begin{quote}
\textit{``Then let the river carry what you are ready to release. Not the name you were given. The weight you have carried because of it.''}
\end{quote}

\vspace{1em}

He untied the braid. Felt the silk of the hair slip through his fingers. Took the red thread and wound it three times around the locket.

The first turn: the way the Cord had taught him to bind a promise.

The second turn: the way Bryn had shown Margot how to tie a knot that would hold.

The third turn: the way Margot had intended to teach her children—a lesson that ended with her own silence before she could pass it on.

He placed the locket on the muddy bank. The water lapped at its edge, testing.

\begin{quote}
\textit{``You have to let go,'' his mother said.}
\end{quote}

\begin{quote}
\textit{``I have let go of everything.''}
\end{quote}

\begin{quote}
\textit{``Not everything. Not yet.''}
\end{quote}

He closed his eyes.

He opened his hand.

The water rose—not in a flood, but in a gentle lift, as if the river itself were a hand taking the locket from his grasp. It slipped beneath the surface with a soft, almost inaudible sigh.

When he looked up, she was gone.

But the mist had thinned. The first true light of dawn spilled over the eastern hills. The Black River ran clear and cold, carrying nothing but light.

\vspace{1em}

He walked back to the garrison. He did not look back. He did not need to.

The gauntlet on his left hand, which had felt like ice when he first rose, now radiated a quiet warmth.

\section*{The Silver Gauntlet}

The gauntlet had always been heavy. Forged in his father's workshop, it had been a symbol of duty, of obligation, of the weight of a name he had never wanted.

But that night, walking back to the garrison, Roran understood something he had not understood before.

The gauntlet was not a chain.

It was a witness.

It had seen every battle he had fought, every oath he had sworn, every blow he had taken for another. It had heard his mother's name spoken over the locket. It had felt the warmth of the braid through his palm.

The gauntlet was his Thiasos.

Not a spirit summoned from the deep. Not a hound bound by oath. A piece of steel, forged in the fires of a father who never loved him, remade in the crucible of a life he had chosen for himself.

It did not speak. It did not need to. It remembered.

\vspace{1em}

Roran sat on the edge of his cot. He removed the gauntlet and laid it on the blanket. The silver was warm.

He traced the engraving along the cuff—the words his father had ordered carved there, a litany of victories and titles. He had never been able to read them without anger.

Now, they were just marks. The ledger was closed.

He thought of Bryn, her brass hook catching the cold light of the courtroom. She had carried her loss for forty years. She had forged it into a tool. She had not let it break her.

He thought of his mother, standing on the water, wrapped in a red blanket. She had not asked for anything. She had simply witnessed.

He thought of the locket, drifting downstream, carrying a braid of auburn hair and a scrap of red thread. It would reach Bryn. Or it would not. The river would decide.

\begin{quote}
\textit{``A debt was paid. Not by me. To me.''}
\end{quote}

He had said those words to the Queen. She had not asked whose.

The gauntlet was warm.

\section*{The Thiasos of the Unbroken Vow}

Roran never spoke of the Black River again. But he carried the gauntlet always. It was not a symbol of his father's expectations. It was a record of his own choices.

When he swore an oath, the gauntlet grew warm.

When he showed mercy, it cooled.

When he stood between an enemy and an innocent, it grew heavy—not with weight, but with presence.

The gauntlet was his Codex. His Rites were engraved on the inside of the cuff, hidden from anyone who did not know to look. His Thiasos was not a separate creature. It was the steel itself. The silver that had witnessed his transformation.

In the years that followed, young knights would ask him how he knew when to strike and when to stay his hand.

He would tap the gauntlet and say:

\begin{quote}
\textit{``This one remembers. It remembers every vow I kept. Every mercy I showed. Every hand I did not raise. The steel does not forget. Neither should you.''}
\end{quote}

\section*{The Thread Holds}

The locket reached Bryn. Or it did not. The story does not say.

What the story says is this:

Roran wore the gauntlet for the rest of his life. He never took a ransom. He never broke an oath. He never forgot the woman who stood on the water, wrapped in a red blanket, and told him to let go.

He died in the Queen's service, old and tired and surrounded by knights who wept at his passing.

They buried him with the gauntlet on his left hand.

The silver was still warm.

\vspace{1em}

\begin{center}
\textit{The thread held. The water carried. The dead smiled. The living walked on.}
\end{center}

\vspace{1em}

\begin{center}
\textit{That is not tragedy. That is grace.}
\end{center}

\section*{The Gauntlet as Thiasos (Mechanical Summary)}

For players who wish to emulate Roran's covenant, the silver gauntlet functions as a unified Codex-Thiasos (see \textit{Codex and Thiasos Unified}).

\begin{itemize}
\item \textbf{Form:} Sentient (unified)—a silver gauntlet worn on the left hand.
\item \textbf{Personality:} Silent, watchful, patient. It does not speak, but it remembers. The Runekeeper knows its temper through warmth, weight, and a faint vibration.
\item \textbf{Witness:} Always counts as a witness for Rites.
\item \textbf{Codex:} Rites are engraved on the inside of the cuff, visible only to the Runekeeper.
\item \textbf{Abilities:}
\begin{itemize}
\item \textit{Warmth of the Vow:} When the Runekeeper swears an oath, the gauntlet grows warm. The Runekeeper gains +1 die to all actions serving that oath.
\item \textit{Cool of Mercy:} When the Runekeeper shows mercy, the gauntlet cools. The Runekeeper gains 1 Boon, usable only for acts of protection or sanctuary.
\item \textit{Weight of Witness:} The gauntlet remembers every blow it has turned. Once per scene, the Runekeeper may spend 1 Boon to reduce incoming Harm by 1 (to a minimum of 0).
\end{itemize}
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
