\section{Talents by Tier}

This table organizes all Runic Warrior talents by tier and category. Talents marked with an asterisk (*) are Patron-specific; refer to the relevant Patron chapter for full descriptions.

\subsection{Foundation Talents (2-4 XP)}

\begin{longtable}{p{0.3\textwidth} p{0.5\textwidth}}
\hline
\textbf{Talent} & \textbf{Effect Summary} \\
\hline
Bonded Focus (4 XP) & Sense unified Thiasos location; summon to hand within Near range. \\
Flesh Codex (2 XP) & Rites tattooed on body; cannot be stolen or confiscated. \\
Iron Stomach (3 XP) & Immune to mundane poisons; halve complications from toxic sources. \\
Living Thiasos (3 XP) & Thiasos gains +1 Cap; acts independently once per scene. \\
Oath-Bound (4 XP) & Reroll failed Body roll when serving a sworn oath. \\
Scarred Resilience (3 XP) & Ignore 1 Fatigue from physical exertion once per scene. \\
The Red Thread (4 XP)* & Tie a knot; allies gain +1 die to one type of roll. \\
Talking Weapon (2 XP) & Sentient Thiasos can speak aloud; counts as witness. \\
The Blade Remembers (4 XP) & Recall past enemy tactics; gain +1 die against familiar foes. \\
Engraved Witness (3 XP) & Codex-Thiasos unified; Rites cannot be destroyed by mundane means. \\
\hline
\end{longtable}

\subsection{Advanced Talents (6-8 XP)}

\begin{longtable}{p{0.3\textwidth} p{0.5\textwidth}}
\hline
\textbf{Talent} & \textbf{Effect Summary} \\
\hline
Anathema to Stagnation (8 XP)* & End infusion early to cancel Harm or Condition; object destroyed. \\
Beast Form (6 XP)* & Partial transformation into a predator for one scene. \\
Binding Presence (6 XP)* & Compel a promise from a target; Resiste test DV 4. \\
Blood Hunt (4 XP)* & +1 die to track wounded targets. \\
Cite Precedent (4 XP)* & Reduce Rite DV by citing a past judgment. \\
Cord of Flesh (8 XP)* & Store Fatigue as visible knots; appear uninjured. \\
Dawn's Judgment (6 XP)* & Extend Rite of Dawn's Judgment for full scene. \\
Engraved Defense (6 XP) & Codex grants +1 Armor; convert Harm to Obligation. \\
Final Witness (6 XP)* & Ask one question about a dying creature's final thoughts. \\
Gavel's Authority (6 XP)* & Strike gavel; hostile NPCs test Resolve or hesitate. \\
Glimpse Beyond (4 XP)* & Intuit hidden detail; gain one piece of information. \\
Harvest Resonance (4 XP)* & After a defeat, gain +1 die or clear 1 Fatigue. \\
The Hilt Listens (6 XP) & Leave Thiasos as listening post; it reports conversations. \\
Iron Resolve (7 XP) & When reduced to 0 Harm, remain standing with 1 Harm and mark 2 Fatigue. \\
Justicar's Immunity (8 XP)* & Cannot be arrested in lands recognizing Mykkiel's law. \\
Lightning Reflexes (4 XP)* & +1 die to defense against ranged attacks. \\
Momentum of the Molt (8 XP)* & After breaking a Condition, next action gains +1 die and improved Position. \\
Perfect Recall (7 XP) & Thiasos recalls witnessed conversations verbatim. \\
Pressure Adaptation (8 XP)* & Immune to underwater penalties; hold breath twice as long. \\
Quiet Step (6 XP)* & +1 die to stealth under moonlight. \\
Rapid Animus (4 XP)* & Create temporary infusion as reaction; +0 Obligation. \\
Rot Sense (6 XP)* & Sense decay, rot, or corruption within Near range. \\
Self-Optimising (6 XP)* & Improve one ability or possession for one scene. \\
Sentinel's Stand (6 XP)* & +1 Armor when holding a threshold; enemies test to cross. \\
Shadow Meld (6 XP)* & Step between shadows within Near range. \\
Shared Burden (6 XP) & Transfer 1 Fatigue to Thiasos; it becomes Winded. \\
Shed Skin (4 XP)* & Escape a Condition restricting movement; mark 1 Fatigue. \\
Shortcut (4 XP)* & Reduce travel clock by 2 segments; GM may impose minor cost. \\
Silver Memory (4 XP)* & Perfect recall of events witnessed under moonlight. \\
Storm Caller (6 XP)* & Peal of thunder; enemies test Resolve or lose minor action. \\
Swift Invocation (6 XP) & Invoke Rite as part of attack or defense action once per scene. \\
Threshold Keep (8 XP)* & Sense active thresholds within Far range; know their mood. \\
Truth's Gaze (6 XP)* & Target cannot lie to you for one minute. \\
Tug the Strand (4 XP)* & Name relationship between NPCs; GM reveals one secret. \\
Unseen Strike (4 XP)* & Attack from hidden position ignores 1 Armor. \\
Ward Sense (4 XP)* & Sense active wards; know their purpose and strength. \\
Witnessed Scar (8 XP) & Touch a scar; gain +1 die on roll tied to that event. \\
\hline
\end{longtable}

\subsection{Prestige Talents (10-12 XP)}

\begin{longtable}{p{0.3\textwidth} p{0.5\textwidth}}
\hline
\textbf{Talent} & \textbf{Effect Summary} \\
\hline
Alpha's Roar (8 XP)* & Terrifying sound; enemies test Resolve or become Shaken. \\
Avatar of the Codex (12 XP) & Become embodiment of Codex; Body +2, Thiasos fights as second combatant. \\
The Bonded Blade (11 XP) & Thiasos performs minor independent actions; warns of danger. \\
The Edge Speaks (8 XP)* & Thiasos speaks in voice of whose blood it has tasted; target may become Shaken. \\
Embraced Corruption (12 XP) & Choose corruption trait; Push Songs without Fatigue once per session. \\
Fractal Response (8 XP)* & +1 die to defense roll; glimpse outcome in advance. \\
Kraken's Warning (8 XP)* & Sense ambush from sea creatures one round in advance. \\
Mercy's Thread (8 XP)* & Tie thread when showing mercy; target gains Promise Clock. \\
The Object Remembers (8 XP) & Unified Thiasos recalls past events; can testify as witness. \\
Reoccupied Vessel (10 XP)* & Raise willing spirit into corpse as Bound Vessel. \\
The Scar That Remembers (11 XP) & Touch a scar; speak sentence as witnessed truth in court. \\
The Spiral's Final Witness (10 XP)* & Interpose between ally and killing blow; Promise Clock fills. \\
The Thiasos Remembers (12 XP) & Thiasos bears witness to past event; testimony considered truth. \\
The Unbroken Vow (10 XP) & Swear oath before witness; gain +2 dice to actions serving it. \\
The Voice of the Fallen (10 XP) & Thiasos speaks in voice of deceased whose blood it has tasted. \\
\hline
\end{longtable}

\subsection{Patron-Specific Talents by Patron}

\subsubsection{Oath of Flame \& Light}
\begin{itemize}
\item Dawn's Judgment (6 XP)
\item Unquenchable (4 XP)
\item The Light Reveals (8 XP)
\end{itemize}

\subsubsection{Oath of Mercy \& Grace}
\begin{itemize}
\item Spiral Counter (6 XP)
\item The Unbuckled Gorget (4 XP)
\item The Spiral's Final Witness (10 XP)
\end{itemize}

\subsubsection{Mykkiel}
\begin{itemize}
\item Cite Precedent (4 XP)
\item Gavel's Authority (6 XP)
\item Justicar's Immunity (8 XP)
\end{itemize}

\subsubsection{Inquisitor Prime}
\begin{itemize}
\item Truth's Gaze (6 XP)
\item Sanctified Weapon (4 XP)
\item Purity Zeal (6 XP)
\end{itemize}

\subsubsection{The Witness}
\begin{itemize}
\item Recorded Echo (4 XP)
\item Truth's Gaze (6 XP)
\item The Scar That Testifies (8 XP)
\end{itemize}

\subsubsection{The Sealed Gate}
\begin{itemize}
\item Ward Sense (4 XP)
\item Sentinel's Stand (6 XP)
\item Threshold Memory (8 XP)
\end{itemize}

\subsubsection{The Traveler}
\begin{itemize}
\item Shortcut (4 XP)
\item Road Memory (6 XP)
\item Wayfinder's Blessing (8 XP)
\end{itemize}

\subsubsection{The Pale Shepherd}
\begin{itemize}
\item Comforting Presence (4 XP)
\item Final Witness (6 XP)
\item Threshold Keep (8 XP)
\end{itemize}

\subsubsection{Grimmir}
\begin{itemize}
\item Seasonal Sense (4 XP)
\item Wild Shape (6 XP)
\item Beast Speech (8 XP)
\end{itemize}

\subsubsection{The Carrion-King}
\begin{itemize}
\item Harvest Resonance (4 XP)
\item Rot Sense (6 XP)
\item Essence Feast (8 XP)
\end{itemize}

\subsubsection{The Savage Heart}
\begin{itemize}
\item Beast Form (6 XP)
\item Blood Hunt (4 XP)
\item Alpha's Roar (8 XP)
\end{itemize}

\subsubsection{Khemesh}
\begin{itemize}
\item Crushing Silence (6 XP)
\item Deep Sight (4 XP)
\item Pressure Adaptation (8 XP)
\end{itemize}

\subsubsection{Xhak'Thul}
\begin{itemize}
\item Storm Caller (6 XP)
\item Lightning Reflexes (4 XP)
\item Fear Eater (8 XP)
\end{itemize}

\subsubsection{Raéyn}
\begin{itemize}
\item Tide-Step (4 XP)
\item Storm Glass (6 XP)
\item Kraken's Warning (8 XP)
\end{itemize}

\subsubsection{Ikasha}
\begin{itemize}
\item Shadow Meld (6 XP)
\item Unseen Strike (4 XP)
\item Raven's Witness (8 XP)
\end{itemize}

\subsubsection{The Silent Choir}
\begin{itemize}
\item Burden Steady (4 XP)
\item Calm Presence (6 XP)
\item Final Vow (10 XP)
\end{itemize}

\subsubsection{Aveh}
\begin{itemize}
\item Storm-Step (6 XP)
\item Forgotten Face (4 XP)
\item Escape Clause (8 XP)
\end{itemize}

\subsubsection{Inaea}
\begin{itemize}
\item Tug the Strand (4 XP)
\item Binding Presence (6 XP)
\item Mercy's Thread (8 XP)
\end{itemize}

\subsubsection{Isoka}
\begin{itemize}
\item Shed Skin (4 XP)
\item Venomous Strike (6 XP)
\item Momentum of the Molt (8 XP)
\end{itemize}

\subsubsection{Malachai}
\begin{itemize}
\item Borrowed Scars (6 XP)
\item Thorn's Memory (4 XP)
\item Cursed Resilience (8 XP)
\end{itemize}

\subsubsection{Clockwork Monad}
\begin{itemize}
\item Rapid Animus (4 XP)
\item Self-Optimising (6 XP)
\item Anathema to Stagnation (8 XP)
\end{itemize}

\subsubsection{Varnek Karn}
\begin{itemize}
\item The Silent Ledger (6 XP)
\item Final Witness (4 XP)
\item Reoccupied Vessel (10 XP)
\end{itemize}

\subsubsection{The High Mother}
\begin{itemize}
\item Cord of Flesh (8 XP)
\item The Open Palm, Closed Cord (8 XP)
\item The Red Thread (4 XP)
\end{itemize}

\subsubsection{The Ninth}
\begin{itemize}
\item Glimpse Beyond (4 XP)
\item Pattern Recognition (6 XP)
\item Fractal Response (8 XP)
\end{itemize}

\subsubsection{Lunara}
\begin{itemize}
\item Silver Memory (4 XP)
\item Quiet Step (6 XP)
\item Seam-Walker (8 XP)
\end{itemize}

\subsubsection{Lethai Syncretic (Ghe'hai)}
\begin{itemize}
\item Mask of Many Faces (6 XP)
\item Silk Binding (4 XP)
\item Ghe'hai's Escape (8 XP)
\end{itemize}

\subsection{Talent Synergies Reference}

\begin{longtable}{p{0.3\textwidth} p{0.6\textwidth}}
\hline
\textbf{Combination} & \textbf{Synergy Effect} \\
\hline
Flesh Codex + Witnessed Scar & Body is both Codex and testimony; cannot be disarmed. \\
Shared Burden + Living Thiasos & You and Thiasos endure together; neither falls alone. \\
Iron Resolve + Avatar of the Codex & Nearly unkillable for one scene; cost is a scar. \\
Talking Weapon + The Blade Remembers & Thiasos is chronicler; reminds you of past tactics. \\
The Hilt Listens + Perfect Recall & Perfect spy; reports conversations verbatim days later. \\
The Edge Speaks + The Voice of the Fallen & Speaks with voices of the dead; confuses and extracts. \\
The Bonded Blade + Engraved Witness & True extension of self; acts in small ways while you focus. \\
\hline
\end{longtable}

The thread held. The water carried. That is not a Rite. That is grace.
