\section{Glossary of Terms}

The following terms are used throughout this expansion. For terms not listed here, consult the \textit{Fate's Edge Comprehensive Resource Guide}.

\subsection*{A}

\paragraph{Bloodsworn}
A Runic Warrior of the High Mother. A warrior who stores Fatigue as knots in their flesh, who pays in blood what others pay in gold.

\subsection*{B}

\paragraph{Runic Warrior}
A Runekeeper whose Codex is a physical object (flesh, stone, or steel) rather than a book. Runic Warriors emphasize physical action, resilience, and close combat.

\paragraph{Bound Vessel}
A corpse reoccupied by a willing spirit through the Rite of Reoccupied Vessel (Varnek Karn). Serves for a set term or until a specific task is completed.

\subsection*{C}

\paragraph{Cairn-Knight}
A Runic Warrior of Varnek Karn. A death-negotiator who fights alongside ancestral shades and carries the weight of the dead into battle.

\paragraph{Codex}
The physical record of a Runekeeper's Rites. For a Runic Warrior, the Codex may be a gorget, shield, staff, tattoo, or sentient weapon.

\paragraph{Collector}
A hollowed child or spirit sent by Malachai to collect debts. Has milk-white eyes sewn shut with silver wire. Does not speak. Watches.

\subsection*{D}

\paragraph{Dawn Warrior}
A Runic Warrior of the Oath of Flame and Light. A crusader who carries vows in steel and fire, who smites foes with radiant strikes.

\paragraph{Death-Warden}
A Runic Warrior of the Pale Shepherd. A warrior who stands at the threshold between life and death, guiding those who cross.

\paragraph{Deep Warrior}
A Runic Warrior of Khemesh. An eldritch knight who fights in crushing darkness, who carries the pressure of the abyss in their lungs.

\subsection*{F}

\paragraph{Fractal Warrior}
A Runic Warrior of The Ninth. A warrior who fights with impossible angles and precognition, whose Codex is a broken mirror.

\subsection*{G}

\paragraph{Gear-Warrior}
A Runic Warrior of the Clockwork Monad. An artificer who wears their craft as armor, who improves themselves mid-battle.

\paragraph{Ghe'hai}
Lethai elite operatives—warriors who are also diplomats, assassins who are also negotiators. A Runic Warrior of the Lethai syncretic tradition.

\paragraph{Grave-Warden}
A Runic Warrior of the Carrion-King. A warrior who harvests endings and turns them into strength, who fights with rot and renewal.

\paragraph{Grey Gauntlet}
A knightly tradition within the Oath of Mercy and Grace. Warriors who carry mercy in steel, who protect the helpless and turn violence back on itself.

\subsection*{H}

\paragraph{Hollowed}
A child or animal whose heart has been removed and replaced with a bundle of grey thread and the ash of a forgotten name. Eyes sewn open with silver wire. Witnesses Rites. Cannot be harmed by mundane means. Associated with the High Mother.

\paragraph{Hospitaller}
A warrior of the Oath of Mercy and Grace. A combat medic who fights not to kill, but to protect. Binds wounds mid-battle. Creates sanctuary.

\subsection*{J}

\paragraph{Justicar}
A Runic Warrior of Mykkiel. A lawbringer who carries the covenant in steel, who enforces oaths with a blade.

\subsection*{P}

\paragraph{Purifier}
A Runic Warrior of the Inquisitor Prime. A witch-hunter who carries judgment in steel and fire, who hunts the corrupt and the tainted.

\subsection*{R}

\paragraph{Road-Warden}
A Runic Warrior of the Traveler. A warrior who carries the road in their feet and their steel, who guides and protects travelers.

\paragraph{Root-Knight}
A Runic Warrior of Grimmir. A wild warrior who fights with root and thorn, who wears the forest as armor.

\subsection*{S}

\paragraph{Serpent-Blade}
A Runic Warrior of Isoka. A warrior who sheds identity like skin, who transforms in the midst of battle, who uses venom as a weapon.

\paragraph{Shadow Warrior}
A Runic Warrior of Ikasha. A ninja, a spy, an unseen blade who strikes once and vanishes.

\paragraph{Silent Blade}
A Runic Warrior of the Silent Choir. A warrior who has sworn an oath of restraint, who protects the vulnerable and keeps the secrets of the dead.

\paragraph{Storm-Sworn}
A Runic Warrior of Xhak'Thul. A primal warrior who carries the tempest in their blood, who shouts as a weapon.

\subsection*{T}

\paragraph{Thiasos}
A fragment of a Patron's attention made manifest. For a Runic Warrior, the Thiasos may be a warhorse, a spirit-hound, a hollowed bird, or a sentient blade.

\paragraph{Thorn-Sworn}
A Runic Warrior of Malachai. A cursed warrior who has accepted power at a price, who carries their affliction like a second skin.

\paragraph{Tide-Knight}
A Runic Warrior of Raéyn. A warrior who fights on rolling decks and slippery shores, who reads the currents of battle as a pilot reads the sea.

\paragraph{Truth-Sworn}
A Runic Warrior of The Witness. A warrior who carries truth in steel and silence, who cannot lie, who exposes deception.

\subsection*{U}

\paragraph{Unbound Blade}
A Runic Warrior of Aveh. A ronin, a wanderer, a warrior who has cut all ties and answers to no master.

\subsection*{W}

\paragraph{Warden}
A Runic Warrior of the Sealed Gate. A sentinel who stands at the boundary between what belongs and what must be kept out.

\paragraph{Web-Warrior}
A Runic Warrior of Inaea. A fighter who binds foes in invisible threads, who controls the battlefield through entanglement and restraint.

\subsection*{Credits}

This glossary was compiled from the \textit{Fate's Edge Comprehensive Resource Guide}, \textit{The Grey Wanderer's Grimoire}, and \textit{The Runic Warrior: Codices of Flesh, Stone, and Steel}.

The thread held. The water carried. That is not a Rite. That is grace.
