\section{Faction Relationship Matrix}

The following tables outline how major factions across Fate's Edge view Runic Warriors. Use these as guidelines for roleplaying, social navigation, and conflict generation.

\subsection*{Faction Stances}

Each faction's stance is rated as:

\begin{itemize}
\item \textbf{Welcome:} Runic Warriors are accepted, even respected. May receive preferential treatment.
\item \textbf{Tolerated:} Runic Warriors are accepted conditionally. Must follow local customs or face consequences.
\item \textbf{Suspicious:} Runic Warriors are watched. May face scrutiny, delayed service, or higher prices.
\item \textbf{Hostile:} Runic Warriors are hunted or actively opposed. May be arrested, attacked, or driven out.
\item \textbf{Varies:} Stance depends on specific Patron, culture, or circumstances.
\end{itemize}

\subsection*{By Patron Family}

\begin{table}[h]
\centering
\small
\begin{tabular}{llllll}
\hline
\textbf{Faction} & \textbf{Vowed} & \textbf{Threshold} & \textbf{Wild} & \textbf{Deep} & \textbf{Shadow} \\
\hline
Chain-Lanterns (Thepyrgos) & Tolerated & Suspicious & Hostile & Hostile & Hostile \\
Spirit-Shield Correctors (Aeler) & Welcome & Welcome & Tolerated & Suspicious & Suspicious \\
Sum-Abjurers (Aelinnel) & Tolerated & Welcome & Tolerated & Tolerated & Tolerated \\
Break-Reins Riders (Ykrul) & Suspicious & Welcome & Welcome & Suspicious & Tolerated \\
Silk Vigil (Lethai) & Tolerated & Welcome & Welcome & Suspicious & Tolerated \\
Shed Vigil (Lethai) & Tolerated & Tolerated & Welcome & Suspicious & Tolerated \\
Wake-Judges (Linn) & Suspicious & Welcome & Welcome & Suspicious & Suspicious \\
Daughters of the Cord & Suspicious & Tolerated & Tolerated & Suspicious & Tolerated \\
Everflame Inquisition (Ecktoria) & Hostile & Hostile & Hostile & Hostile & Hostile \\
Queen's Justicars (Viterra) & Welcome & Tolerated & Suspicious & Suspicious & Suspicious \\
Admiralty (Kahfagia) & Tolerated & Welcome & Tolerated & Tolerated & Tolerated \\
Salt Prince's Court (Zakov) & Tolerated & Tolerated & Suspicious & Welcome & Tolerated \\
Vermilion Corps (Oshiira) & Tolerated & Suspicious & Suspicious & Suspicious & Hostile \\
Veiled Sisters (Ashaan) & Hostile & Hostile & Hostile & Tolerated & Hostile \\
Green Council (Sekogo) & Suspicious & Tolerated & Welcome & Suspicious & Tolerated \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\small
\begin{tabular}{lllll}
\hline
\textbf{Faction} & \textbf{Web} & \textbf{Forge} & \textbf{Grave} & \textbf{Fracture} \\
\hline
Chain-Lanterns (Thepyrgos) & Hostile & Tolerated & Hostile & Hostile \\
Spirit-Shield Correctors (Aeler) & Suspicious & Welcome & Tolerated & Suspicious \\
Sum-Abjurers (Aelinnel) & Tolerated & Welcome & Tolerated & Welcome \\
Break-Reins Riders (Ykrul) & Tolerated & Tolerated & Welcome & Suspicious \\
Silk Vigil (Lethai) & Welcome & Tolerated & Tolerated & Tolerated \\
Shed Vigil (Lethai) & Tolerated & Suspicious & Tolerated & Suspicious \\
Wake-Judges (Linn) & Suspicious & Suspicious & Welcome & Suspicious \\
Daughters of the Cord & Welcome & Suspicious & Tolerated & Suspicious \\
Everflame Inquisition (Ecktoria) & Hostile & Hostile & Hostile & Hostile \\
Queen's Justicars (Viterra) & Suspicious & Tolerated & Suspicious & Suspicious \\
Admiralty (Kahfagia) & Tolerated & Tolerated & Tolerated & Suspicious \\
Salt Prince's Court (Zakov) & Tolerated & Tolerated & Welcome & Suspicious \\
Vermilion Corps (Oshiira) & Hostile & Tolerated & Suspicious & Hostile \\
Veiled Sisters (Ashaan) & Hostile & Suspicious & Hostile & Hostile \\
Green Council (Sekogo) & Welcome & Suspicious & Welcome & Suspicious \\
\hline
\end{tabular}
\end{table}

\subsection*{By Specific Patron (Notable Examples)}

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\hline
\textbf{Patron} & \textbf{Faction} & \textbf{Stance} \\
\hline
Oath of Flame \& Light & Everflame Inquisition & Welcome (as censer-knights) \\
Oath of Flame \& Light & Queen's Justicars & Welcome (as dawn-knights) \\
Oath of Mercy \& Grace & Daughters of the Cord & Tolerated (rivals in mercy) \\
Mykkiel & Queen's Justicars & Welcome (as justicars) \\
Inquisitor Prime & Chain-Lanterns & Welcome (as witch-hunters) \\
The Sealed Gate & Spirit-Shield Correctors & Welcome (as wardens) \\
The Traveler & Break-Reins Riders & Welcome (as road-wardens) \\
Grimmir & Green Council & Welcome (as root-keepers) \\
The Savage Heart & Ykrul clans & Welcome (as beast-warriors) \\
Khemesh & Salt Prince's Court & Welcome (as deep-warriors) \\
Xhak'Thul & Ykrul clans & Welcome (as storm-sworn) \\
Ikasha & Daughters of the Cord & Tolerated (shadow-kin) \\
Aveh & Break-Reins Riders & Tolerated (unbound blades) \\
Inaea & Silk Vigil & Welcome (as web-warriors) \\
The High Mother & Daughters of the Cord & Welcome (as bloodsworn) \\
Clockwork Monad & Spirit-Shield Correctors & Welcome (as artificers) \\
Varnek Karn & Wake-Judges & Welcome (as cairn-knights) \\
\hline
\end{tabular}
\end{table}

\subsection*{Using the Faction Matrix}

When a Runic Warrior interacts with a faction, consider:

\begin{itemize}
\item \textbf{Welcome:} NPCs may offer aid, lower prices, or share information. Social rolls start at Controlled Position.
\item \textbf{Tolerated:} NPCs are neutral but watchful. Social rolls start at Controlled Position but may shift to Desperate if the Runekeeper draws attention.
\item \textbf{Suspicious:} NPCs are wary. Social rolls start at Controlled Position but suffer +1 DV. Guards may ask questions.
\item \textbf{Hostile:} NPCs are actively opposed. Social rolls start at Desperate Position. The faction may send agents to harass, arrest, or eliminate the Runekeeper.
\item \textbf{Varies:} The GM determines stance based on specific Patron, cultural context, and recent events.
\end{itemize}

\subsection*{Changing Stance}

A Runic Warrior's standing with a faction can change through:

\begin{itemize}
\item \textbf{Service:} Completing a mission for the faction improves stance by one step (Hostile → Suspicious → Tolerated → Welcome).
\item \textbf{Harm:} Attacking faction members or undermining their goals worsens stance by one step.
\item \textbf{Reputation:} Famous (or infamous) deeds may shift stance region-wide.
\item \textbf{Patron's Influence:} Some Patrons have standing relationships with factions (e.g., the Everflame Inquisition welcomes Oath of Flame \& Light Runekeepers).
\end{itemize}

\subsection*{Faction Clocks}

For campaigns involving significant faction interaction, use the following clocks:

\begin{itemize}
\item \textbf{Faction Attention [6]:} Ticks when the party attracts notice (successful missions, public failures, high-profile conflicts). When full, a faction representative makes contact—ally, employer, or adversary.
\item \textbf{Faction Wrath [8]:} Ticks when the party harms faction interests. When full, the faction sends agents to eliminate or capture the party.
\item \textbf{Faction Favor [4]:} Ticks when the party aids faction interests. When full, the party may call in a significant favor (safe passage, resources, intelligence).
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
