\section{The Weave's Receipt}

Every spell is a transaction. The Weave provides the effect; the caster provides the intention. But the Weave also issues a receipt---a record of the transaction's friction. We call it \emph{backlash}.

Students fear backlash. They imagine it as punishment, a lash, a scar. It is none of these things. Backlash is \emph{color}. A spark that did not exist before. A chill that passes through the room. A witness who shivers and does not know why. The Weave does not punish. It \emph{remarks}.

I have seen a student weep at a minor backlash. I have seen another shrug off a catastrophic failure and ask for more chalk. The difference is not in the Weave. It is in the caster's relationship with imperfection.

\subsection{On the Nature of Backlash}

Backlash occurs when the Weave is asked to do something it found \emph{interesting}. A clean success with no ones produces no backlash because the Weave had no comment. A success with ones produces minor backlash---a comment, a footnote, a raised eyebrow. A partial success produces moderate backlash---a paragraph of unsolicited advice. A miss produces major or catastrophic backlash---a full critique, delivered in the language of consequence.

The severity of backlash is also influenced by the caster's Position when they cast. A Dominant casting---controlled, prepared, witnessed---insulates the caster from the worst of the Weave's commentary. A Desperate casting---rushed, frightened, improvisational---invites the Weave to speak more freely.

I have cast in Dominant Position many times. The Weave was polite. I have cast in Desperate Position twice. The Weave was \emph{memorable}.

\subsection{The Spectrum of Receipts}

\begin{longtable}{p{0.2\textwidth} p{0.35\textwidth} p{0.35\textwidth}}
\caption{Backlash Severity and Its Meaning}\\
\midrule
\textbf{Severity} & \textbf{What the Weave says} & \textbf{What it feels like}\\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Severity} & \textbf{What the Weave says} & \textbf{What it feels like}\\
\midrule
\endhead
Minor & \emph{``Noted.''} & A spark, a chill, a whisper. A single point of Fatigue. A distraction that costs a die. The Weave is acknowledging the transaction. \\
Moderate & \emph{``Explain yourself.''} & The spell works, but the Weave adds a condition---Blinded, Silenced, Entangled. A ward collapses. An alarm sounds. The Weave is asking for better grammar. \\
Major & \emph{``We need to talk.''} & Harm. A permanent scar. An unwanted transformation. The spell inverts---healing becomes harm, a veil becomes a beacon. The Weave is correcting the record. \\
Catastrophic & \emph{``I will remember this.''} & A reality fracture. A Patron manifests. A threshold opens where none should exist. The caster's name is added to a ledger that never closes. The Weave is taking a personal interest. \\
\hline
\end{longtable}

\subsection{Backlash by Element}

The flavor of backlash depends on the element of the spell. A [BURNING] spell that backfires produces smoke and sparks. A [WAVE] spell produces a slick floor and damp clothes. The Weave is consistent in its inconsistency.

\begin{longtable}{p{0.2\textwidth} p{0.35\textwidth} p{0.35\textwidth}}
\caption{Elemental Backlash Examples}\\
\midrule
\textbf{Element} & \textbf{Minor (1--2 SB)} & \textbf{Moderate (3--4 SB)}\\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Element} & \textbf{Minor (1--2 SB)} & \textbf{Moderate (3--4 SB)}\\
\midrule
\endhead
Fire & Smoke, sparks, a singed sleeve & An uncontrolled flame ignites nearby objects \\
Water & A damp floor, a splash, soggy papers & A small flood pools around the caster's feet \\
Earth & A tremor, dust in the eyes, a cracked flagstone & A binding root grips the caster's ankle \\
Air & A sudden gust, a deafening ring, scattered papers & A strong wind pushes the caster off balance \\
Fate & A door closes unexpectedly, a coin lands on edge & A prophecy---the Weave speaks a future that may be avoided \\
Life & A patch of grass grows through the floor, a nosebleed & A fever---the caster shivers and sweats for an hour \\
Luck & A small object falls from a shelf & A friend suffers a minor misfortune nearby \\
Death & A chill, a whisper of a forgotten name & A ghostly presence watches for a scene \\
\hline
\end{longtable}

\subsection{A Lesson in Humility}

I was examining a student in her third year. She was talented---perhaps the most talented I had taught in a decade. She attempted a [GATE] during a practical examination, in Desperate Position, because she had miscalculated the time and the examination hall was about to close.

She rolled a miss.

The portal opened. It did not lead to the exit. It led to a basement storage room filled with old examination papers. The papers flew out like startled birds. The Weave added a commentary: every paper that landed on the floor had a failing grade marked in red ink that was not there before.

The student was not harmed. She was, however, required to reshelve the entire archive. She passed the examination the following year. She never cast [GATE] in Desperate Position again.

The Weave's receipt is not a punishment. It is a \emph{reminder} of what you owe to preparation.

\subsection{When Backlash is a Gift}

I have learned to welcome minor backlash. A spark that lights a candle you needed lit. A chill that alerts you to a spirit in the room. A whisper that contains a word you were missing.

The Weave is not your enemy. It is your conversation partner. Sometimes it interrupts. Sometimes the interruption is the point.

A student once cast [SCRY] to locate a stolen artifact. She rolled a partial success and earned a minor backlash. The Weave showed her not the artifact's location, but a reflection of the thief's face in a puddle. She did not find the artifact that day. She found the thief. The artifact followed.

She thanked the Weave. I have never seen a student thank the Weave before. The Weave did not respond. It rarely does. But the next spell she cast was a clean success.

Coincidence? Perhaps. Perhaps not.

\subsection{A Final Observation on Receipts}

Keep a record of your backlashes. I do. I have a ledger labeled \emph{``Interest Paid.''} It contains a thousand entries: \emph{``Burnt sleeve, 862 AR.''} \emph{``Frog in throat, 864 AR.''} \emph{``Portal to the wrong basement, 867 AR.''}

The Weave does not charge interest. It charges \emph{attention}. A caster who ignores backlash is a caster who will eventually owe more than they can pay. A caster who studies backlash is a caster who learns the Weave's sense of humour.

And the Weave, I have learned, has an excellent sense of humour. It is merely very, very dry.

\textit{--- Lysandra}
