\section{The Small Work}

Every master was once a student who could not light a candle without scorching the wick. I have scorched hundreds. The Weave forgives incompetence in small doses; it reserves its wrath for ambition without foundation.

Cantrips are spells of one tag. They are the alphabet of the Weave---individual letters, not yet words. A student who masters cantrips may still fail at greater workings, but a student who scorns cantrips will fail spectacularly. I have seen it. I have graded it.

\subsection{On the Nature of Cantrips}

A cantrip requires one tag, one intention, and a moment of focus. The Difficulty Value is two---the Weave's lowest threshold. A competent caster with a pool of four or five dice will succeed more often than not. A novice will learn what failure tastes like. Both will improve.

The following table lists cantrips that have survived centuries of student misuse. I present them without apology for their modesty. A modest spell that works is worth a dozen ambitious spells that set fire to the syllabus.

\medskip
\noindent
\begin{longtable}{p{0.2\textwidth} p{0.15\textwidth} p{0.5\textwidth}}
\caption{Common Cantrips of the Humble Hand}\\
\midrule
\textbf{Spell} & \textbf{Tag} & \textbf{What It Does (in practice)}\\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Spell} & \textbf{Tag} & \textbf{What It Does (in practice)}\\
\midrule
\endhead
Ember Flick & [BURNING] & Ignites a small, unattended flammable object within Close range. A candle, a twist of paper, a dry leaf. Not a person. The Weave draws a line at arson. \\
Hush & [SILENCE] & Mutes all sound within a Close zone for a few heartbeats. A dropped key, a muttered curse, a footstep on a creaking stair---the Weave swallows them. The caster's own voice is not muted. Some lessons are experiential. \\
Warmth & [HEAL] & Removes one point of Fatigue from the caster or a touched ally. The Weave does not knit flesh; it merely reminds the body that it is tired. A gentle spell. Often underestimated. \\
Folded Sight & [VEIL] & Blurs the caster's features for a single exchange. Not invisibility. The observer sees a tradesman, a servant, a face they will not remember. The effect ends if the caster speaks their own name aloud. \\
Knot of Stillness & [BIND] & Holds a single small object in place---a door, a latch, a coin on a table. The Weave does not lock; it suggests. A determined child could overcome it. A distracted guard will not notice. \\
Pale Light & [LIGHT] & Illuminates the caster's hand with a soft, steady glow. The light reaches Near range but dims before Far. Cannot be used to blind. Cannot be used to start fires. The Weave is specific. \\
Whisper & [MESSAGE] & Sends a sentence of up to eight words to a willing target within Near range. The target hears the words in the caster's voice, but no one else does. The Weave does not deliver lies. It delivers sounds. \\
\hline
\end{longtable}

\subsection{On Failure, Minor}

A cantrip's backlash is modest because the cantrip's ambition is modest. The Weave does not punish a stutter; it merely repeats the word incorrectly.

\noindent
\begin{tabular}{p{0.25\textwidth} p{0.7\textwidth}}
\textbf{Ember Flick (Miss)} & The spark jumps back. The caster's sleeve smoulders. Mark one Fatigue. \\
\textbf{Hush (Partial)} & The sound muffles but does not vanish. A witness tilts their head. The GM gains a Story Beat. \\
\textbf{Warmth (Miss)} & The Fatigue remains. The caster feels a phantom chill instead. The Weave has misread the prescription. \\
\textbf{Folded Sight (Partial)} & The caster's features blur, but a distinctive scar or piece of jewelry remains visible. The witness may recognize the detail later. \\
\end{tabular}

\subsection{The First Spell I Learned}

I was seventeen. My instructor, a woman whose name I have forgotten but whose patience I have not, placed a candle on a stone and told me to light it with my mind. I failed for three days. On the fourth day, I lit the candle. On the fifth day, I lit my sleeve.

The spell was [BURNING]. The tag was one. The difficulty was two. I rolled a miss and earned two boons, which I spent immediately to re-roll the die that had come up as a one. The second roll succeeded. The candle lit. The Weave did not applaud. My instructor nodded once and said, \emph{``Again.''}

That is the nature of cantrips. They are not impressive. They are foundations. Build them well, and the walls will follow. Build them poorly, and the walls will collapse. I have seen both.

\subsection{When to Use a Cantrip}

Use a cantrip when the cost of failure is manageable. A [LIGHT] that flickers is an inconvenience. A [GATE] that misfolds is a tragedy.

Use a cantrip when you are learning. The Weave is a patient teacher, but it expects homework.

Use a cantrip when you are tired. A cantrip's backlash will not kill you. It may embarrass you. It may singe your sleeve. It will not open a door to the Hollow.

Save the dangerous tags for when the candle is already burning.

\subsection{A Final Observation}

I have seen a master of Spellcraft light a candle with a glance. I have seen the same master extinguish a charging berserker's will with a whispered [FEAR]. Both are one-tag spells. Both rely on the same grammar. The difference is not in the tag. The difference is in the caster's understanding of when silence is a weapon.

Practice the small work. The large work will follow.

\textit{--- Lysandra}
