\section{The Dangerous Door}

The Synod calls them ``unstable quanta.'' The Chain−Lanterns call them ``proscribed manifestations.'' I call them \emph{interesting}. A dangerous tag is not forbidden. It is merely expensive. The Weave adds a surcharge—not because it disapproves, but because folding reality requires more ink than lighting a candle.

These spells are the ones that rewrite the scene. A [LEAP] that bypasses a locked door. A [GATE] that connects two rooms that are not adjacent. A [TRANSFORM] that changes a guard into a frog. A [DOMINATE] that makes an enemy lay down their sword. The Weave permits these effects. It also \emph{watches} them more closely.

I have cast dangerous spells. I have survived most of them. The ones I did not survive, I do not discuss.

\subsection{On the Nature of Dangerous Tags}

A dangerous tag adds a surcharge of two to the spell's Difficulty instead of one. The Weave does not explain why. I have my theories. Reality has inertia; persuading it to fold, teleport, or alter its substance requires more energy than persuading it to warm your hands.

The tags in this chapter are:

\begin{itemize}
    ℹtem \textbf{[LEAP]} — The caster moves from one location to another without crossing the space between. Short range. Line of sight required. The Weave does not like being asked to ignore distance.
    ℹtem \textbf{[FOLD]} — Similar to [LEAP], but the caster may bend space around an obstacle. A wall, a door, a guard's line of sight. The Weave dislikes being asked to ignore geometry.
    ℹtem \textbf{[GATE]} — A two− way portal between two locations. Range can be considerable. The Weave dislikes being asked to create permanent exceptions.
    ℹtem \textbf{[TRANSFORM]} — The caster changes a target's form. A person into a newt, a sword into a ribbon, a lock into a puddle of wax. The Weave dislikes being asked to forget what a thing is.
    ℹtem \textbf{[CREATE]} — The caster manifests a mundane object from nothing. A key, a rope, a coin. The object lasts for a scene, then crumbles to dust. The Weave dislikes being asked to pay for purchases it did not approve.
    ℹtem \textbf{[SUMMON]} — The caster calls a spirit, elemental, or minor Outsider into their presence. The creature is not compelled to obey. Negotiation follows. The Weave dislikes being asked to open doors that should remain closed.
    ℹtem \textbf{[DOMINATE]} — The caster imposes their will on a sentient creature. The target resists. The Weave dislikes being asked to erase agency.
    ℹtem \textbf{[REVIVE]} — The caster reverses time for a few heartbeats. An action is undone. A wound is unmade. The Weave \emph{hates} being asked to contradict itself.
\end{itemize}

\subsection{Common Dangerous Spells}

\noindent
\begin{longtable}{p{0.2\textwidth} p{0.28\textwidth} p{0.42\textwidth}}
∩tion{Spells of the Dangerous Door}\\
∣rule
\textbf{Spell} & \textbf{Tags} & \textbf{What It Does (in practice)}\\
∣rule
\endfirsthead
μlticolumn{3}{c}{\tablename\ θble{} −− continued}\\
∣rule
\textbf{Spell} & \textbf{Tags} & \textbf{What It Does (in practice)}\\
∣rule
\endhead
Threshold Step & [LEAP] & The caster vanishes and reappears anywhere within Near range that they can see. A wall is still a wall. A closed door is still a door. The Weave does not give you eyes on the other side. \\
Fold the Wall & [FOLD][AREA] & The caster bends space around a single barrier—a wall, a hedge, a line of guards. They and anyone holding their hand step through to the other side. The Weave complains about the extra passengers. \\
Open the Way & [GATE] & The caster creates a shimmering portal between two locations they have visited. The portal lasts for three heartbeats. The Weave charges a toll in memory—a fact, a name, a recent conversation. You will not notice the loss until you need it. \\
New Skin & [TRANSFORM][SELF] & The caster alters their own appearance. Face, build, voice—all change. The new face is not a specific person's; it is a plausible stranger. The Weave does not forge documents. \\
Turn the Blade & [TRANSFORM][STRIKE] & The caster touches a weapon. The weapon becomes a harmless object—a flower, a ribbon, a dead fish. The effect lasts for a few heartbeats. The Weave's sense of humour is evident in the result. \\
Coin from Ash & [CREATE] & The caster clenches their fist and opens it to reveal a single silver coin. The coin is real for one scene, then crumbles. The Weave does not care about inflation. \\
Call the Wisp & [SUMMON] & The caster whistles a specific note. A minor spirit—a wisp, a shade, a house− brownie—appears within Near range. The spirit is not a servant. It may be bargained with. The Weave does not mediate. \\
Speak My Will & [DOMINATE][TARGET] & The caster speaks a single imperative—\emph{``Stop.''} \emph{``Kneel.''} \emph{``Sleep.''}—and the target must resist or obey. The Weave does not care about the morality of the command. \\
Breath Returned & [REVIVE][TOUCH] & The caster touches a creature that has died within the last few heartbeats. The Weave reverses time locally. The creature lives. Something else dies in their place. The Weave does not explain the accounting. \\
ℎline
\end{longtable}

\subsection{On Failure, Dangerous}

A dangerous spell that fails is not a mishap. It is a \emph{conversation} with a reality that is now paying attention.

\begin{tabular}{p{0.25\textwidth} p{0.7\textwidth}}
\textbf{Threshold Step (Miss)} & The caster vanishes and reappears in the same spot. The Weave has noted their destination and found it unacceptable. The caster marks one Fatigue. The GM gains two Story Beats. \\
\textbf{Open the Way (Partial)} & The portal opens, but it leads to a different location—similar, but wrong. The room is furnished differently. The person waiting is not an ally. The GM chooses the destination. \\
\textbf{New Skin (Miss)} & The caster's appearance changes, but the change is unstable. Their face shifts every few heartbeats. Strangers are unsettled. Allies are confused. The effect lasts until the caster sleeps. \\
\textbf{Turn the Blade (Partial)} & The weapon transforms, but only halfway. A sword becomes a sword− shaped ribbon. It still cuts, but it also flaps in the breeze. The target may still use it, at a penalty. \\
\textbf{Call the Wisp (Miss)} & The spirit arrives. It is not the spirit the caster called. It is larger. It is hungrier. It is not hostile—yet. But it is \emph{interested}. The GM gains three Story Beats. \\
\textbf{Speak My Will (Partial)} & The target resists the command, but the effort leaves them disoriented. They lose their next action. They also remember that someone tried to dominate them. \\
\textbf{Breath Returned (Any outcome)} & Something dies. If the spell succeeds, the Weave takes a life of equivalent weight—a bird, a rat, a child in a distant village. If the spell fails, the Weave takes the caster's memory of the person they tried to save. The Weave does not refund exchanges. \\
\end{tabular}

\subsection{The Lecture I Do Not Give Anymore}

I used to teach a seminar on dangerous tags. I stopped after the incident with the gate.

A student—bright, impatient, from a family of minor nobility—attempted [GATE] during a supervised examination. He had performed the spell three times in practice. The fourth time, the Weave answered differently.

The portal opened. It did not lead to the adjacent hallway. It led to a room I have never seen, filled with mirrors facing each other. In the infinite reflections, the student saw versions of himself that had made different choices. One was a murderer. One was a saint. One was a corpse.

He closed the portal. He sat down. He did not speak for three days. When he finally spoke, he asked to withdraw from the University. He is now a sheep farmer in Acasia. He writes me once a year. His letters do not mention magic.

The Weave is not a toy. It is a language. Dangerous tags are the profanity. Use them when you must. Use them when the alternative is worse. Do not use them to impress your peers.

\subsection{When to Use Dangerous Tags}

Use [LEAP] when the door is locked and the guard is turning.

Use [GATE] when you are fleeing and the road behind you is on fire.

Use [TRANSFORM] when the enemy's weapon is the only thing keeping you from the exit.

Use [DOMINATE] when a friend is about to make a lethal mistake and will not listen to reason.

Do not use [REVIVE]. The Weave's accounting is cruel. I have seen it. I will not describe it.

\subsection{A Final Observation on the Door}

Every dangerous spell opens a door. Some doors close when you walk through them. Some doors stay open, waiting for something to follow. The Weave does not close doors it did not open. That is your responsibility.

Cast with intention. Cast with necessity. Cast with a plan for what comes back with you.

\textit{−−− Lysandra}
