\section{The Elemental Chorus}

Fire, water, earth, air. The old tetrad. Every student learns them in the first week; every master learns to fear them in the first decade. The Weave does not favor one element over another. It favors \emph{clarity}. A fire that knows its shape is safer than a breeze that does not know its direction.

These spells are the loud ones. They leave scorch marks, puddles, cracked stones, and ringing ears. The Synod notices loud spells. So do witnesses. Cast with appropriate discretion.

\subsection{On the Nature of Elemental Magic}

An elemental spell persuades the Weave to borrow a quality from the physical world. [BURNING] is not fire; it is the \emph{idea} of fire, given temporary citizenship in reality. [WAVE] is not water; it is the \emph{memory} of water, shaped by the caster's will.

The Weave is not a reservoir of elements. It is a grammar of \emph{resonance}. You speak [STORM], and the Weave hums in sympathy with every lightning strike within a mile. You speak [STONE], and the Weave recalls the weight of a mountain it has never seen but can imagine.

The tags in this chapter are drawn from the eight elements recognized by the Thepyrgos curriculum. I list them here for reference, though the spells that follow will demonstrate their use.

\begin{itemize}
    \item \textbf{Fire:} [BURNING], [STORM] (lightning), [FLARE]
    \item \textbf{Water:} [WAVE], [FLOOD], [MIST]
    \item \textbf{Earth:} [STONE], [ROOT], [DUST]
    \item \textbf{Air:} [WIND], [GUST], [THUNDER]
    \item \textbf{Fate:} [CHANCE], [OMEN], [KNOT]
    \item \textbf{Life:} [GROWTH], [BLOOM], [WITHER]
    \item \textbf{Luck:} [FORTUNE], [MISFORTUNE], [BALANCE]
    \item \textbf{Death:} [ASH], [REST], [SILENCE] (in its terminal form)
\end{itemize}

Some elements are safer than others. Fire is predictable. Fate is not. I advise students to master the tetrad before reaching toward the higher notes.

\subsection{Common Elemental Spells}

\noindent
\begin{longtable}{p{0.2\textwidth} p{0.28\textwidth} p{0.42\textwidth}}
\caption{Spells of the Elemental Chorus}\\
\midrule
\textbf{Spell} & \textbf{Tags} & \textbf{What It Does (in practice)}\\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Spell} & \textbf{Tags} & \textbf{What It Does (in practice)}\\
\midrule
\endhead
Ember Dart & [BURNING][STRIKE] & A bolt of concentrated heat flies from the caster's hand toward a single target within Near range. It sears cloth, singes skin, and spoils uncovered food. Lethal? No. Distracting? Very. \\
Rising Mist & [WATER][AREA] & A thick, cool fog billows from a point the caster chooses, filling a Near zone within Near range. Vision beyond a few paces is impossible. The fog persists for a few heartbeats, then dissipates. The Weave does not appreciate stasis. \\
Stone Grip & [EARTH][BIND] & Stone or packed earth beneath a target's feet softens, then hardens. The target's ankles are held fast. A strong pull may break the bond. A wise target does not struggle. \\
Sudden Gale & [AIR][FORCE] & A blast of wind erupts from the caster's palm, pushing a single target one range band away from the caster. The target may keep their feet if they are expecting the shove. The Weave does not guarantee grace. \\
Storm Spark & [STORM][STRIKE] & A thin line of lightning jumps from the caster's finger to a metal object within Near range. The object becomes briefly hot. A creature holding the object marks one Fatigue. The crack of thunder is audible for a considerable distance. \\
Rippling Step & [WATER][LEAP] & The caster dissolves into a spray of droplets and reforms within Near range at a location they can see. Passing through grates or under doors is possible if the gap is not sealed. The Weave does not clean the caster's clothes. \\
Dust Veil & [EARTH][VEIL] & A cloud of fine dust rises around the caster, obscuring their exact position. Attacks against the caster suffer uncertainty. The dust settles quickly. The caster's cough is not magical. \\
Cinder Breath & [BURNING][AREA] & The caster exhales a cone of hot ash and cinders. Everyone in Close range before them must shield their eyes or suffer the distraction. The caster's throat will be sore for an hour. \\
Tremor Tap & [EARTH][WAVE] & The caster strikes the ground with their foot or a staff. A ripple of motion travels through the earth, knocking creatures in a Close line off balance. The Weave amplifies the tap but does not aim. \\
\hline
\end{longtable}

\subsection{On Failure, Elemental}

Elemental failure is rarely subtle. The Weave does not misplace a syllable; it \emph{overstates} the resonance.

\begin{tabular}{p{0.25\textwidth} p{0.7\textwidth}}
\textbf{Ember Dart (Miss)} & The bolt flares wide, igniting a nearby object. A curtain, a stack of papers, a guard's sleeve. The GM gains a Story Beat and may tick a Fire Hazard timer. \\
\textbf{Rising Mist (Partial)} & The fog rises, but it also drifts toward the caster. Visibility for everyone is reduced, including allies. The caster's own breath fogs before their eyes. \\
\textbf{Stone Grip (Miss)} & The stone softens, but it does not harden again. The floor becomes unstable. Anyone stepping on that spot within the next few minutes may stumble. The GM gains a Story Beat. \\
\textbf{Sudden Gale (Partial)} & The wind pushes the target, but the caster is also pushed backward one step. The Weave has shared the momentum. Neither party is where they intended to be. \\
\textbf{Storm Spark (Miss)} & The lightning does not strike the intended target. It strikes the nearest metal object---which may be the caster's own belt buckle. Mark one Fatigue. The thunder is still loud. \\
\textbf{Dust Veil (Partial)} & The dust rises, but it also coats the caster's eyes and mouth. They cannot see clearly for a moment. Their next action suffers a penalty. \\
\end{tabular}

\subsection{A Lesson from the Quadrangle}

A student in my Elemental Theory seminar---a young man from the Vilikari marches, eager and careless---attempted [STORM][STRIKE] during a practical examination. The target was a bronze disc on a stand. The tags were two. The Difficulty was three.

He rolled a miss.

The lightning did not strike the disc. It struck the iron chandelier above the examination hall. The chandelier fell. No one was injured, but the noise attracted a Synod clerk, who wrote a report, which found its way to the Chain- Lanterns, who visited my classroom the following week.

The student passed the course. He did not pass the year. He transferred to a college that studied agricultural thaumaturgy, where he now grows larger potatoes with [GROWTH] and [BLOOM]. I am told he is content.

The Weave does not punish ambition. It merely \emph{redirects} it.

\subsection{When to Use the Elemental Chorus}

Use fire when you need a distraction. Use water when you need to not be seen. Use earth when you need someone to stay where they are. Use air when you need someone to go somewhere else.

Use storm when you are prepared to be noticed.

Do not use the higher elements---Fate, Luck, Life, Death---unless you have mastered the tetrad. A mispronounced [BURNING] is a singed eyebrow. A mispronounced [FORTUNE] is a decade of bad luck. The Weave's interest rates are higher for abstract nouns.

\subsection{A Final Observation on the Tetrad}

The old masters taught that the four elements are the foundation of all magic. They were wrong. The foundation of all magic is \emph{intention}. The elements are merely the vocabulary.

Still, a rich vocabulary is useful. You can intimidate a guard with a wall of fire. You cannot intimidate them with a well- intentioned pause.

Learn the elements. Then learn when to remain silent.

\textit{--- Lysandra}
