\chapter{Tag Reference}

The Weave speaks in tags. These are its vocabulary. A student who knows the tags but not their combinations is like a merchant who knows every coin but cannot count. A student who knows the combinations but not the tags is a poet who has forgotten the alphabet.

The following table lists the tags I have found most useful in four decades of casting. It is not exhaustive. The Weave adds new words occasionally, usually when a caster attempts something the Synod has not yet named. When that happens, write to me. I will update the syllabus.

\section{How to Read This Reference}

Each entry includes the tag's name, its family, and a plain description of what it does. I have avoided numbers. Numbers are for clerks. The Weave understands intent, not arithmetic.

A tag marked with a dagger (†) is considered \emph{dangerous}. The Weave charges a surcharge for these tags—not because they are forbidden, but because they require the Weave to fold, teleport, or unmake something. The surcharge is not a punishment. It is a recognition of difficulty.

\section{Tag Reference Table}

\noindent
\begin{longtable}{p{0.2\textwidth} p{0.15\textwidth} p{0.55\textwidth}}
\caption{Tags of the Weave}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{What It Does}\\
\midrule
\endfirsthead
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\midrule
\textbf{Tag} & \textbf{Family} & \textbf{What It Does}\\
\midrule
\endhead
BURNING & Elemental (Fire) & Ignites, heats, or smokes a target. A candle, a torch, a dry leaf. Not a person. The Weave draws a line. \\
FREEZING & Elemental (Water) & Cools, chills, or encases in frost. A lock may stiffen. A puddle may harden. A breath may fog. \\
STORM & Elemental (Air) & Calls lightning, thunder, or static charge. Loud. Visible. The Weave does not offer discreet lightning. \\
STONE & Elemental (Earth) & Hardens, weights, or shapes stone and earth. A wall may rise. A floor may crack. A pebble may become a brick. \\
WAVE & Elemental (Water) & Moves water or creates pressure. A current may push. A wave may crash. A cistern may empty. \\
WIND & Elemental (Air) & Moves air. A gust may shove. A breeze may cool. A vacuum may gasp. \\
FORCE & Kinetic & Pushes, pulls, or strikes with invisible pressure. A door may slam. A person may stagger. A weapon may fly from a hand. \\
AREA & Kinetic & Expands a spell's effect to cover a zone. Everyone within Near is affected. The Weave does not distinguish friend from foe. \\
STRIKE & Kinetic & Focuses a spell on a single target. Precision over spread. The Weave appreciates economy. \\
WALL & Kinetic & Creates a barrier of force, element, or illusion. A wall may block. A wall may hide. A wall may fall. \\
BIND & Kinetic & Restrains a target. Roots, ropes, or invisible chains. The restraint is temporary. The Weave does not imprison. \\
DISPEL & Kinetic & Unravels an ongoing magical effect. A ward may drop. A veil may lift. A curse may pause. \\
VEIL & Mind & Conceals a target from sight, memory, or attention. Not invisibility. \emph{Forgettability}. \\
SCRY & Mind & Reveals a distant location or object. A glimpse, not a full view. The Weave does not offer tours. \\
MEMORY & Mind & Touches the past. A recollection may sharpen, blur, or vanish. The Weave charges interest. \\
COMMAND & Mind & Compels a single short action. \emph{``Stop.''} \emph{``Kneel.''} \emph{``Drop it.''} The Weave does not write essays. \\
FEAR & Mind & Instills dread in a target. They may flee, freeze, or falter. The Weave does not dictate which. \\
HEAL & Life & Restores Fatigue. Mends minor wounds. The Weave is a physician, not a miracle worker. \\
PURIFY & Life & Removes poison, disease, or corruption. The Weave cleans what it can. Some stains are permanent. \\
STRENGTHEN & Life & Temporarily enhances a body. A faster sprint, a harder punch, a steadier hand. The Weave lends, not gives. \\
WAKEN & Life & Rouses a sleeping or stunned creature. The Weave does not care if they wanted to wake. \\
BEAST & Life & Communicates with or influences an animal. A horse may calm. A dog may track. A rat may deliver a message. \\
LEAP† & Space & Teleports the caster within Near range. Line of sight required. The Weave does not give you eyes on the far side. \\
FOLD† & Space & Bends space around an obstacle. A wall, a door, a guard. The Weave dislikes geometry. \\
GATE† & Space & Creates a two- way portal between two locations. The Weave charges a toll in memory. \\
GRAVITY† & Space & Alters local gravity. A target may float, stumble, or crush. The Weave is not gentle. \\
CREATE† & Creation & Manifests a mundane object from nothing. A key, a rope, a coin. The object crumbles after a scene. \\
SUMMON† & Creation & Calls a spirit, elemental, or minor Outsider. The creature is not compelled. Negotiation follows. \\
TRANSMUTE† & Creation & Changes one substance into another. Lead to gold, stone to water, air to glass. The change is temporary. \\
ANIMATE† & Creation & Gives motion to an inanimate object. A rope may tie itself. A statue may turn its head. A broom may sweep. \\
TRANSFORM† & Creation & Changes a creature's form. A guard into a newt, a sword into a ribbon. The Weave dislikes forgetting. \\
DOMINATE† & Mind & Imposes the caster's will on a sentient creature. The target resists. The Weave does not mediate. \\
REWIND† & Time & Reverses a few heartbeats. An action is undone. A wound is unmade. The Weave hates this. \\
\hline
\caption*{Dangerous tags are marked with †. The Weave charges a surcharge for their use. The surcharge is not negotiable.}
\end{longtable}

\section{On Dangerous Tags}

A student once asked me, \emph{``If dangerous tags are so costly, why use them at all?''}

I replied, \emph{``Because sometimes the door is locked, the guard is turning, and the alternative is worse.''}

Dangerous tags are not for daily practice. They are not for examinations. They are not for impressing your peers in the tavern. They are for the moments when a clean success is the difference between walking away and being carried.

Use them sparingly. Use them deliberately. Use them with a clear intention and an exit strategy.

And when you use them, accept the receipt. The Weave will remember. So should you.

\section{A Final Note on Missing Tags}

You may notice that certain tags—[TIME], [FATE], [TRUTH]—appear in other grimoires but not in this reference. I have omitted them deliberately. They are not more powerful than the tags listed here. They are merely more dangerous. A caster who has mastered the tetrad, the unseen hand, and the defensive arts may eventually reach for these higher notes. That caster no longer needs this book.

Until then, trust the vocabulary you have. A well- spoken [BURNING] is worth a dozen mumbling [FATES].

\textit{--- Lysandra}
