% Fate's Edge --- System Reference Document (SRD)
% Main LaTeX file to compile the full document.
\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\tcbuselibrary{breakable, skins}

\geometry{
  letterpaper,
  margin=1in
}

\setlength{\parskip}{0.6em}
\setlength{\parindent}{0pt}
\setlist[itemize]{topsep=0.2em,itemsep=0.2em}
\setlist[description]{style=nextline}

\title{\vspace{-1.5em}Fate's Edge\\\large System Reference Document}
\author{}
\date{\today}

\begin{document}
\maketitle
\newpage

\chapter*{Copyright and License}

\section*{Fate's Edge: Essential Guide}

Copyright \copyright\ Nicholas A. Gasper

This document, \textit{Fate's Edge: Essential Guide}, is licensed under the Creative Commons Attribution 4.0 International License. To view a copy of this license, visit:

\texttt{http://creativecommons.org/licenses/by/4.0/}

You are free to:
\begin{itemize}[leftmargin=*]
\item \textbf{Share} --- copy and redistribute the material in any medium or format.
\item \textbf{Adapt} --- remix, transform, and build upon the material for any purpose, even commercially.
\end{itemize}

Under the following terms:
\begin{itemize}[leftmargin=*]
\item \textbf{Attribution} --- You must give appropriate credit, provide a link to the license, and indicate if changes were made. You may do so in any reasonable manner, but not in any way that suggests the licensor endorses you or your use.
\end{itemize}

\section*{System Reference Document (SRD)}

The \textit{Fate's Edge System Reference Document} is also licensed under the Creative Commons Attribution 4.0 International License, under the same terms as above.

\section*{Commercial Products}

The following products are \textbf{not} covered by the CC-BY license and are available for purchase:

\begin{itemize}[leftmargin=*]
\item \textit{Fate's Edge: Player's Guide}
\item \textit{Fate's Edge: GM Guide}
\item \textit{Fate's Edge: The Amaranthine Core}
\item \textit{Fate's Edge: Adventures and Expansions}
\end{itemize}

These works are the intellectual property of Nicholas A. Gasper and may not be reproduced, distributed, or transmitted in any form or by any means without prior written permission, except for brief quotations in reviews and certain other noncommercial uses permitted by copyright law.

For permission requests, contact the author at the address provided in the commercial editions.

\section*{The Author's Intent}

The Essential Guide and SRD are offered freely to lower the barrier to entry, to allow community contributions, and to support third-party creators. The commercial products fund further development, art, and the continued existence of Fate's Edge as a living game. Thank you for your support.
\newpage
\tableofcontents
\newpage

% --- PART 1: FOUNDATION ---
\input{sections/00-quickstart}           % 2-page quickstart (new)
\input{sections/01-core-principles}      % safety, tone, central question
\input{sections/02-core-mechanic}        % resolution loop, DV, Position, Outcome, Boons, SB

% --- PART 2: CHARACTERS ---
\input{sections/03-character-creation}   % step-by-step, bonds, complications
\input{sections/04-attributes-skills}    % merged from old 1.2 + 13.1
\input{sections/05-talents}              % all talents (incl. summoning, cantor)
\input{sections/06-equipment}            % weapons (Harm by weight), armor, gear, condition

% --- PART 3: ACTION & COMBAT ---
\input{sections/07-action-economy}       % turn structure, Move/Action, reactions
\input{sections/08-combat}               % range, cover, stealth, mounted
\input{sections/09-health-fatigue-harm}  % Fatigue, Harm, armor conversion, recovery

% --- PART 4: MAGIC & SUPERNATURAL ---
\input{sections/10-magic-system}         % four paths, Obligation, Corruption, Backlash
\input{sections/11-tags-effects}         % element tags, effect tags, DV calculation
\input{sections/12-rites-invokers}       % universal rituals, invoker rituals
\input{sections/13-summons}              % Call → Bind → Leash, spirit templates

% --- PART 5: HOLDINGS & ADVANCEMENT ---
\input{sections/14-followers-assets}     % merged (was 5 & 6)
\input{sections/15-advancement-xp}       % XP awards, spending, reputation tiers

% --- PART 6: TRAVEL & ENCOUNTERS ---
\input{sections/16-travel-framework}     % legs, roles, timers, supply
\input{sections/17-encounters}           % social, environmental, magical (merged)
\input{sections/18-complications-timers} % SB menu, timers, hazards

% --- PART 7: OPTIONAL SYSTEMS ---
\input{sections/19-deck-consequences}    % optional
\input{sections/20-crown-spread}         % optional campaign frame
\input{sections/21-player-archetypes}    % optional (Solo/Mixed/Mastermind)

% --- PART 8: GM TOOLKIT (new) ---
\input{sections/22-gm-guidelines}        % consolidated GM advice, session zero
\input{sections/23-npc-templates}        % NEW: bandits, guards, spirits, beasts

% --- APPENDICES ---
\input{sections/appendix-quickref}       % all tables in one place
\input{sections/appendix-example-spells} % expanded spell/rite examples
\input{sections/appendix-example-characters} % 2-3 pregens

\end{document}
