\chapter{The Vent That Whispered}

\section*{Told by}
Old Orra, lamp‑keeper of the Third Stack, Aelerian Range. Not been underground in thirty years, but still counts her breaths before sleep.

\section*{The Tale}

You ever walk the Deep Galleries alone? No? Good. Don't start.

The air down there has a ledger. Eight breaths is a corridor. Sixteen is a chamber. Thirty‑two is a fool who forgot to turn back. That's what they taught me. Breathe, count, turn. Never answer the ninth.

I was young. Lamp trimmed, chalk dry. Third Stack to the old sump. Easy walk. Should've been two hours.

At the junction -- you know the one, three tunnels, the middle one always drips -- I stopped. Something was wrong. The drip had stopped. The echo had changed. And then I heard it.

Not with my ears. Inside. Like a memory that wasn't mine.

\emph{"Count the ninth breath."}

I didn't. I walked. The voice followed.

\emph{"The ninth holds the name of the miner who died here. The year the pillar cracked. The route to the lost sump."}

My lamp flickered. I didn't stop to trim it. I kept counting. Eight in. Eight out. My lips moved. The voice couldn't get in.

When I reached the surface, the lamp was dry. My shadow was gone. The next morning I looked in a mirror and the reflection smiled first.

Took a week for the shadow to come back. Took another for my left hand to stop writing in a script I couldn't read. The Vent‑Prior said the deep had spoken and I hadn't answered, and that was mercy.

She promoted me to lamp‑keeper. I never went down again.

\section*{The Witch's Closing}
"Stone remembers. So does the space between breaths. Answer neither. And if you must walk the deep, count to eight -- then start over."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Voice in the Crack}

\paragraph{Description:}
A psychic remnant of a Deep Drake sealed in the mountain. It cannot move, but it whispers into the minds of those who walk alone in deep, dark, air‑starved places. It offers secrets. It asks for the ninth breath.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item Tags: [PSIONIC] [DOMAIN (stone)] [DREAD]
\item Harm: none directly -- mental strain.
\item Fatigue: On failed Spirit + Resolve (DV 4), mark 1 Fatigue and hear a false memory.
\item Clock: \textit{Counting the Ninth} [4] -- ticks each time a character answers a question while alone in the dark. When full, lose shadow for 1d3 days (social Position --1; shadow may act independently).
\item Strings: The Voice can offer a \textit{True Shem} -- a forgotten fact, lost name, or route through stone -- in exchange for counting the ninth breath (advancing the timer).
\end{itemize}

\paragraph{Motivation:}
Feeds on attention and isolation. Wants to be witnessed.

\paragraph{Adventure Hook:}
A vent‑warden has gone missing in the Third Stack. The party finds a crack in the wall -- and hears breathing on the other side.

\section*{Colophon}
Recorded in the Third Stack hostel, winter of 862, by Dusana of the Raven Road. The chalk circle was erased before I slept. My shadow arrived a heartbeat late.

