\chapter{The Hollow Bride}

\section*{Told by}
Mabs the Miller, a round woman with flour on her apron and a cider‑cup always half full. Tells stories while she works the grindstone. Says the Hollow talks to her at night.

\section*{The Tale}

You know about the ninth cup, don't you? Course you do. Every child knows: you pour eight. You never pour the ninth. The ninth is for the Hollow.

But children forget. Or they get curious. That's how it always starts.

Little Kesta -- she was what, seven? -- went to the well at dusk. The adults were at the harvest feast. She was supposed to draw water for the morning bread. She counted the buckets: one, two, three, four, five, six, seven, eight.

Then she thought: what's one more?

She drew the ninth bucket. The water was black. And something climbed out.

It wore her mother's wedding dress. It smiled with her mother's mouth. But its eyes were the colour of a drowned moon.

"Hollow Bride," Kesta whispered. That's what the old folk call it. You don't name it. She named it.

The Bride followed her home. It stood outside the red door. It knocked. Three knocks, then a pause, then three more. The rhythm of a heart that stopped.

Kesta's mother opened the door. "Come in," she said. "You must be cold." That's the law of hospitality. A red door means bread, salt, and one safe night.

The Hollow Bride stepped inside. It sat at the table. It didn't eat. It didn't speak. It just watched.

At midnight, it said: "You poured the ninth cup. You named me. Now I am your guest."

The mother understood. She cut a loaf with a knife that had been blessed at the Moot Oak. She put salt on the table with her left hand. She said: "Guest‑right lasts one night. At dawn, you leave."

The Hollow Bride smiled. "I will leave," it said. "But I will return. The ninth cup is an invitation. The threshold remembers."

Kesta is an old woman now. She still sets eight cups on her table. She still leaves the ninth empty. And every harvest, when the lanterns burn blue at the ford, she checks the red door. So far, it's held.

\section*{The Witch's Closing}
"Hospitality is a thread. Tie it right, and it binds the guest. Tie it wrong, and it binds you. And never, ever pour the ninth."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Hollow Bride}

\paragraph{Description:}
A threshold‑spirit that wears the face of a loved one. It can only enter a home if invited -- either by name, by pouring a ninth cup, or by breaking some other taboo of hospitality.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [FAE] [THRESHOLD] [DREAD]
\item Cannot cross a red door unless invited. Repelled by bread, salt, and a blessed knife.
\item Harm: none. It imposes Conditions: \textit{Watched}, \textit{Invited}, \textit{Named}.
\item Clock: \textit{The Ninth Cup} [6] -- ticks each time hospitality is broken. When full, the Bride can enter uninvited.
\item Bargaining: The Bride may be offered a substitute -- a memory, a name, a firstborn child -- to buy another year of silence.
\end{itemize}

\paragraph{Adventure Hook:}
A village has the same nightmare: a woman in a wedding dress outside a red door, knocking. The party must find who broke guest‑right and HEAL the threshold.

\section*{Colophon}
Recorded in the miller's kitchen, year of the sour harvest. The kettle was warm. The cider was dry. The Bride did not knock that night.

