\chapter{The Warrior Who Refused to Call the Runner}

\section*{Told by}
Temur Many‑Scars, steppe rider with a braid down to his belt and a laugh like breaking bones. Been in more camps than he can count. Says the only thing that kept him alive was bad geometry.

\section*{A Note on the Telling}
The Ykrul speak of the board with their own tongue. \textit{War‑Chief} is the piece others call Blue. \textit{Runner} is Green. \textit{Well} is Sanctum. \textit{Calling a Runner} is Seeding. \textit{Nailed} is Rooted. \textit{The Five‑Breath Race} is Reforge. \textit{Cuts} are specials. \textit{Breaths} are plies. \textit{The shadow of the stone} is ZoC.

\section*{The Tale}

The Khagan called a muster. We were young, hungry, and stupid. My cousin -- bold as a winter wolf -- wanted glory. He pushed his War‑Chief forward. He didn't count exits. He just pushed.

The enemy stepped aside. Let him pass. Let him think he was winning.

By the time he reached the Well, his cuts were spent. He looked at the far Well. It was empty. He could Call a Runner. But his War‑Chief would be Nailed for a breath.

I said: "Don't. Their War‑Chief still has a cut. They're two breaths away."

He laughed. "They won't risk it."

He Called.

The enemy hopped a stone, landed on the Well, and displaced his War‑Chief off the board. Five breaths started ticking. My cousin had to plant his banner in five of his breaths.

He didn't. The enemy had already welded the approach. Every lane ended in the shadow of the stone. The race was dead.

We lost. The clan paid tribute for a year.

The Khagan demoted him to herd‑watcher. "You forgot the first number," she said. "Count exits before you count glory. Call only when the geometry is yours."

\section*{The Witch's Closing}
"A free War‑Chief is a noose. Call only when the shadow is yours. Always keep one cut for the door."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Herd‑Watcher's Regret (a curse)}

\paragraph{Description:}
A punishment for warriors who Call unsafely. The curse takes reputation, rank, sense of direction. The victim becomes forgettable -- still alive but no longer counted.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [SOCIAL] [DREAD]
\item --1 Position on Command and Sway.
\item Once per session, GM may force a re‑roll on Navigation or Tactics (wrong path).
\item Strings: Curse lifted by a public act of heroic geometry -- winning a Kon'reh match or completing a Five‑Breath Race.
\item Clock: \textit{Forgotten Name} [6] -- when full, the victim loses their name entirely. No one remembers them.
\end{itemize}

\paragraph{Adventure Hook:}
A decorated warrior has disappeared from the camp ledger. She still sits by the fire, but everyone looks past her. Restore her name.

\section*{Colophon}
Winter camp of the Grey Fox clan. The rider spat tea leaves into the fire. "That was the winter I learned to count."

