\chapter{The Silk Witness}

\section*{Told by}
Maer of the Root‑Circle, a graying wood elf who mends bridges and remembers every broken oath. Speaks like water over stones – soft, but sharp.

\section*{The Tale}

A merchant came to the Valewood, full of himself and his ledgers. He needed a bridge. He had coin. He did not have courtesy.

He spoke to the Lethai‑al wardens. “Name your price,” he said. “Double if you build it by autumn.”

They sent for the Silk Vigil. The Vigil hung a thread across the river. White silk, thin as a spider’s. “Cross this,” they said, “and we will talk.”

The merchant laughed and stepped over.

The thread broke.

The Vigil nodded. “You have refused witness. Build your bridge. Every plank will remember your rudeness. No one will sing its name.”

He built it. It held. Workers called it “the Rudeness.” Travellers spat on its railings.

Years later, he returned, bent and grey. He knelt. “What must I pay?”

The Vigil picked up a single strand. “Say: ‘I am sorry for the courtesy I broke.’ Then build a garden under the bridge, for the eels to hide in.”

He did. The eels came. The bridge softened. Now they sing of it – not as a curse, but as a repair.

\section*{The Witch's Closing}
“Mercy with memory. That is the thread. Cut it, and you build alone. Mend it, and the road belongs to everyone.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Unwitnessed Bridge}

A bridge built without proper rites – no thread, no witness. Sound, but radiates a low social curse. Those who cross feel irritation, suspicion.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [LOCATION] [SOCIAL]
\item –1 Position on social rolls on or near the bridge.
\item Once per day, GM may force a re‑roll on Command or Sway while crossing.
\item Clock: \textit{Repair the Bridge} [6] – needs an offering of courtesy (bread, salt, an apology) and a public act of repair.
\end{itemize}

\paragraph{Adventure Hook:}
A new bridge divides a river community. Merchants whisper of bad luck. The builder died without apologising. Mend the crossing.

\section*{Colophon}
Sat under a roof‑tree at the edge of the Root‑Circle. The eels were listening.

