\chapter{The Gloss That Ate a Name}

\section*{Told by}
Archivist Venn, a high elf with ink under her nails and a voice that never rises above a murmur. Says she’s forgotten more languages than most people learn.

\section*{The Tale}

In the great library of Thepyrgos, there is a scroll that rewrites itself. Copied from a copy from a copy, each margin a correction, each correction a new error.

A scholar – Mira – wanted the original. Archivists refused. “The original is lost.”

She found a forbidden room. Inside, a single scroll sealed with wax not warm for a thousand years. She broke the seal.

The scroll was blank.

Weeks passed. Mira began to forget her own name. Students’ faces blurred. Mother’s voice became a hum. She returned. The scroll now held one word: *Mira*.

She understood. The scroll did not store knowledge. It *ate* it. The copy was safety. The original was hunger.

She burned it. The fire was cold. The smoke smelled of forgotten birthdays. Her name returned – but not her grandmother’s face. That was the toll.

\section*{The Witch's Closing}
“Context is a key. Lose the key, lose the door. Lose the door, lose yourself.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Hungry Scroll}

Ancient artifact that erases memory. Appears as a blank parchment in a sealed case. Reader loses specific memories – names, faces, skills.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [KNOWLEDGE] [DREAD]
\item On reading, test Spirit + Resolve (DV 5). Failure: lose one memory (player chooses a minor fact).
\item Clock: \textit{Erasure} [8]. When full, victim forgets own name and cannot form new long‑term memories.
\item Destroyed only with fire from a living witness (a bell, a lamp, a human breath).
\end{itemize}

\paragraph{Adventure Hook:}
A scholar has gone missing. Her notes are contradictions. In her study, a blank scroll on a lectern. Read it, burn it, or seal it – knowing someone else will find it.

\section*{Colophon}
Copied from a margin note in a banned book. The archivist would not let me see her face. The ink was still wet.

