\chapter{The Bargain of the Shed Skin}

\section*{Told by}
Serpent‑keeper Ashai, a dark elf with scales tattooed up her arms and a voice like a knife wrapped in silk. Says she’s shed three names already.

\section*{The Tale}

A woman came to the Coil Court. “I gave a Lethai‑ar my fear of the dark. Now I cannot sleep. Faces in the shadows not my own.”

The serpent‑queen nodded. “You paid for a shedding. You received it. The price is not your fear. It is your solitude.”

The woman wept. “I did not know.”

“You did not ask.”

She offered a second bargain. “Shed this solitude. Forget the faces. Forget your mother’s lullaby. Owe a favour to the Coil Court.”

The woman agreed. She shed. She slept. A year later, a messenger came. “A child has been born without a shadow. Nurse it until the shadow returns.”

She nursed for seven years. The shadow never returned. The child grew strange – calm, watchful, always polite. When the mother died, the child shed its own name and joined the Coil Court.

The woman was free. But she often looked at her hands and wondered whose heartbeat she was counting.

\section*{The Witch's Closing}
“A shedding does not erase the debt. It only changes the skin.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Shadowless Child}

A child born without a shadow – an old bargain unpaid. Unnervingly calm, immune to fear, can walk thresholds that bar others. May be a vessel for something older.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [FAE] [THRESHOLD] [DREAD]
\item Immune to fear. Cannot be compelled by emotion‑targeting magic.
\item –1 Position on Stealth near the child (shadows are thinner).
\item Cannot be harmed by any creature that casts a shadow (convert Harm to Fatigue).
\item Clock: \textit{Return the Shadow} [8] – find the original bargain and either fulfil or break it.
\end{itemize}

\paragraph{Adventure Hook:}
A village is terrorised by a child who never sleeps, never blinks, and knows everyone’s secrets. The child has no shadow – taken by a Lethai‑ar who now wants it back.

\section*{Colophon}
Recorded in a Coil Court antechamber, with warm milk and bitter herbs. The serpent‑queen did not look at me once.

