\chapter{The Selkie's Oath}

\section*{Told by}
Captain Rana of the Eel‑Skiff, a Linn woman with salt‑bleached braids and a laugh like waves on shingle. Never goes ashore without a knife.

\section*{The Tale}

A selkie married a fisherman. She gave the usual bargain: hide the skin, she stays. Lose it, she leaves.

He hid it under the hearthstone. Ten years, children, mended nets, quiet storms.

Then the child fell ill. Blood in the cough. The nearest healer was a day's sail. The fisherman hesitated. If he left, the skin unguarded. If he stayed, the child died.

The selkie reached under the hearthstone, pulled out the skin, and cut it into strips.

"Bandages," she said. "The child's breath is worth more than my freedom."

She wrapped the child's chest. The fever broke. The skin was ruined.

At dawn, the fisherman asked, "Will you leave?"

She touched his cheek. "I have no skin to return to. But I have a new oath: I choose to stay."

They say she grew webbing between her fingers the next winter, and her eyes turned the colour of the deep. But she stayed.

\section*{The Witch's Closing}
"Oaths change. So do the skins we wear. The sea doesn"t care which -- only that you choose."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Cut Skin}

A ruined selkie skin, now a patchwork of bandages. Heals any wound, but transfers selkie nature to the user -- webbed fingers, taste for raw fish, longing for the sea.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [HEAL] [CURSE]
\item Heals 1 Harm per use. User marks 1 Fatigue after each use (the sea calls).
\item Clock: \textit{Sea‑Change} [4]. When full, user gains a permanent minor mutation (webbed digits, salt‑caked skin, moon‑lit eyes). Cannot be removed without a full renaming ritual.
\end{itemize}

\paragraph{Adventure Hook:}
A healer's bandages cure any sickness -- but her patients dream of drowned cities. The source: a cut selkie skin. Destroy it or keep the miracle.

\section*{Colophon}
Told on the deck of the Eel‑Skiff, rain coming. The captain spat over the rail. "That's the truth."

