\chapter{The Peddler's Swerve}

\section*{Told by}
Rusti the Factor, a Vilikari trader with a missing finger and a smile that never reaches her eyes. She has sold the same horse three times and still owns it. She says that is not theft. That is velocity.

\section*{The Tale}

A peddler came to a border ford. The toll was high. She had no coin. She had a cart full of empty barrels — iron-hooped, tar-sealed, the kind that once held brandy from the southern islands.

"Air," she told the toll-keeper. "Fresh mountain air. The valley pays handsomely. They say the fog down there is thick enough to drown in. My customers prefer to breathe."

He laughed — a thick, wet sound. "Air? I will let you pass for a taste."

She uncorked a barrel with ceremony. Held it to his nose. Pine and snow and the particular cold that lives in high places, the kind that makes you remember a childhood you never had.

He waved her through, still laughing.

The next day, a rival told him the barrels were empty. Furious — not at the loss, at the insult — he rode after her.

She had already sold the air.

Not as air. As "breathing lessons for the rich." Copper per lungful. Copper per desperate merchant who could not afford another winter of valley lung. The barrels were full of gold now, and the air was gone.

He cornered her at the next ford. "You cheated me."

She did not flinch. "Air is free. The mountain gives it to anyone who climbs. You chose not to sell it. I did. That is not cheating. That is seeing what is in front of your face."

She paid double the original toll — from the gold — and asked, "Do you want a lesson in breathing?"

He did not answer. But the next year, he sold fog to the lowlands. Called it "cloud silk." Made a fortune.

The peddler? She kept one barrel empty. Just in case.

\section*{The Witch's Closing}
"Sell the hour, not the horse. Value is not in the thing. It is in the need. And need is a thing you can put in a barrel if you name it correctly."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Hollow Barrel}

A barrel that appears empty but can be filled with anything people believe is inside. It works by suggestion — open it, describe the contents with enough detail, and it produces faint scent, sound, warmth, the ghost-touch of what is not there. It cannot create matter. Only perception. Only the space between what is and what is wanted.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [ILLUSION] [TRICK] [VEIL]
\item Once per scene, a character can declare that the barrel contains a plausible small object — a lock of hair, a poison, a letter, a breath of spice air. The barrel produces sensory evidence: smell, faint sound, a wisp of warmth. A Perception test at DV 3 reveals the illusion. At DV 5, the character notices that the \textit{lack} of weight is wrong.
\item Cannot create fire, light, or anything that would cause physical harm. The barrel is a liar, not a weapon.
\item Used to fool guards, start rumors, create false trails, or convince a desperate person that hope is closer than it is.
\item Clock: \textit{The Barrel's Limit} [6]. Ticks each time the barrel creates an effect in a new location. When full, the barrel cracks. It can still be used, but anyone within Near range hears a faint whisper: \textit{empty}. The jig is up.
\end{itemize}

\paragraph{Adventure Hook:}
A con artist is selling "dragon's breath" and "ghost tears" using hollow barrels. She has made enemies of three guilds and one actual ghost. Catch her — or join her. The ghost is offering better terms.

\section*{Colophon}
Recorded in a Vilikari waystation, bad wine and good lies. The factor winked at me — a slow, deliberate gesture — and pointed at my notebook with the hand missing its smallest finger.

"You will sell that one day," she said.

I have not. Yet.