\chapter{The Censor's Mask}

\section*{Told by}
Old Censor Varo, a man with one eye and a brand on his cheek. He sits by the forum fountain and feed the pigeons. Children dare each other to ask him about the mask.

\section*{The Tale}

The Censor of Ecktoria wears a mask. It is old, bronze, with eyeholes that see through lies. He wears it only at the Triumph, when petitioners approach.

One year, a merchant offered a bribe. The Censor refused. The merchant insisted. The Censor removed the mask.

The merchant screamed. The mask had no face inside. Only a whisper: "I have seen your ledgers."

The merchant fled. The Censor put the mask back on. At the next Triumph, he wore it again.

But he dreamed that night of his own lies -- the grain he had skimmed, the witness he had ignored. The mask had seen him too.

He tried to remove it. It would not come off. He cut the straps. It stayed. He broke the bronze. It reformed.

For a year he wore the mask, unable to eat or drink, but did not starve. The mask fed on his guilt.

On the next Triumph, he knelt before the assembly. "I am unworthy," he said. "Judge me."

The mask fell off. His face was younger. His eye was whole.

He became the most honest Censor the city ever had. He never wore the mask again. He did not need to.

\section*{The Witch's Closing}
"Judgment begins at home. The mask only shows what you already know."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Censor's Mask}

A bronze mask that reveals the wearer's own hidden crimes. It cannot be removed until the wearer confesses publicly. While worn, the wearer cannot eat or drink but does not die -- sustained by guilt.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [TRUTH]
\item Wearer gains +2 dice to Insight against deception. But at the end of each scene, they must confess a secret flaw (GM may ask "what guilt does the mask show you?"). On refusal, timer \textit{Confession} [4] ticks.
\item When timer fills, the wearer is compelled to make a public confession. After confession, the mask can be removed.
\end{itemize}

\paragraph{Adventure Hook:}
A city official has stopped eating. He speaks in whispers, and a bronze mask has fused to his face. The party must learn what he did -- and persuade him to confess before he wastes away.

\section*{Colophon}
Forum fountain, late winter. The old man threw a crust to a pigeon. "I was his clerk," he said. "I forged the ledgers. He wore the mask for both of us."

