\chapter{The Bride of Nine Knots}

\section*{Told by}
Lady Ysoria, a Vhasian widow who lives in a red tower and receives visitors only at dusk. Her left hand has six fingers. She says the extra one was a wedding gift.

\section*{The Tale}

A bride tied her dowry rope with nine knots. One for each year she waited for her betrothed. The priest said, "Eight is custom. Ninth is a weapon."

She tied it anyway.

At the wedding, her rival -- who had also loved the groom -- stepped forward and cut the ninth knot.

The groom forgot his vows. He stood silent, mouth open, eyes vacant. The bride spoke for him: "I give this ring. I take this name. I close this door."

The priest nodded. The bond held.

But the rival's hand withered where she had cut the cord. The bride kept the severed knot in a locket. On her deathbed, she gave it to her daughter.

"If you ever need to silence a liar," she said, "cut a knot. But be sure you know whose tongue will still speak afterwards."

\section*{The Witch's Closing}
"The ninth knot is the thread that holds. Cut it, and the story unravels. But the blade cuts both ways."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Ninth Knot}

A piece of wedding cord that can be used to "un‑speak" a single vow or promise. When cut, the target forgets the specific promise they made. The cutter's hand withers for a year (cannot hold weapons or tools without pain). The knot regrows after a season.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [TRUTH]
\item As an action, cut the knot and name a vow. The target loses all memory of that vow. The cutter suffers --1 die on all physical rolls using their dominant hand for 1 month (in‑game time).
\item The knot can be used once per season.
\end{itemize}

\paragraph{Adventure Hook:}
A noble has been forced into a marriage contract she does not want. She possesses a ninth knot -- but cutting it will also destroy the trust her family has in her. The party must decide: cut the knot, or find another way.

\section*{Colophon}
Red tower, dusk. The widow poured wine into eight cups. The ninth she emptied on the floor. "For the ones who forgot," she said.

