\chapter{The Gate That Asked a Question}

\section*{Told by}
Old Hedge‑warden Cormac, a man who leans on a staff and speaks only in rulings. He has judged more boundary disputes than the queen.

\section*{The Tale}

A traveller came to a gate. It was old stone, moss‑grown, with a single bell hanging from the lintel.

The gate‑keeper sat on a stool, whittling a stick. “What is the weight of a road?” he asked.

The traveller thought. “The weight of the stone beneath it.”

The gate‑keeper shook his head. The bell did not ring.

The traveller tried again. “The weight of the coin paid for it.”

The bell remained silent.

A third answer: “The weight of the promise that keeps it open.”

The bell rang once. The gate opened.

The gate‑keeper said, “A road has no weight. It is permission. Permission can be withdrawn. So be courteous, or walk elsewhere.”

The traveller passed. Behind him, the gate‑keeper erased a chalk mark from the stone. A boundary had moved.

\section*{The Witch's Closing}
“Law is not stone. It is consent. Consent can be withdrawn. So speak softly, and keep your promises.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Wandering Gate}

A gate that appears only at twilight, at different boundaries. It asks a riddle of each traveller. Wrong answers: it disappears. Right answer: the traveller may cross, and the gate moves a boundary stone one inch in the direction the traveller came from – slowly reshaping property lines.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [THRESHOLD] [MYSTERY]
\item To find the gate, a character must be lost at dusk and succeed on a Wisdom + Survival test (DV 4). The gate appears.
\item The riddle is always a moral question about boundaries, property, or promises. The GM may ask the player to answer in character. No roll – the answer must be judged by the GM as “true” or “false” based on the character’s past actions.
\item If correct, the gate opens and may move a boundary stone. If incorrect, the character cannot cross that threshold for a year.
\end{itemize}

\paragraph{Adventure Hook:}
A farmer’s field has shrunk overnight. His neighbour claims the boundary moved. The party must find the wandering gate and answer its riddle before the dispute becomes a blood feud.

\section*{Colophon}
At a hedge‑moot, twilight. The old warden pointed a stick at a chalk line. “That’s where I moved it,” he said. “He never noticed.”

