\chapter{The Stair That Was Not}

\section*{Told by}
Scholar Lyra, who never leaves the library and reads by candlelight. Her handwriting is too small to read without a lens. She says that is intentional.

\section*{The Tale}

A student climbed a stair in the Collegium of Keys. He had counted the steps: one, two, three, four, five, six, seven, eight.

Then the ninth. There was no ninth. The stair continued anyway.

He climbed until his legs ached. The walls changed from stone to vellum to glass. The steps became sentences. The banisters became arguments.

At the top, a desk. On it, a single page. It was his own thesis -- corrected.

He read the corrections. They were right. He had been wrong. He wept.

When he looked up, the desk was gone. The stair was gone. He was standing in the basement, holding a piece of chalk.

He never told anyone what the corrections said. But his next thesis was perfect. And he never climbed a stair without counting twice.

\section*{The Witch's Closing}
"A proof that cannot be improved is already dead. The stair judges. Accept its edits, or stay in the basement."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Invisible Stair}

A stair that appears only to those who have made a significant intellectual error. It leads to a room where their mistake is made manifest. To leave, they must admit the error aloud. Denial traps them until they speak true.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [TRUTH] [DREAM]
\item Trigger: a character fails a Lore or Investigation roll by 5 or more. The stair appears in the next room they enter.
\item To climb, they must succeed a Will + Lore test (DV 6). Failure: they climb but cannot reach the top until they solve the puzzle (GM presents a riddle based on the failed skill check). After solving, they gain +1 die to all future rolls of that skill (insight from the correction).
\end{itemize}

\paragraph{Adventure Hook:}
A scholar's maps have been wrong for decades. She has hidden her error by avoiding the region. The party must find the invisible stair, correct the map, and face whatever the truth has done to the people who lived there.

\section*{Colophon}
Scriptorium of the Collegium, midnight. The scholar held up a lens. "See the margin?" she said. "That's where the stair begins."

