\chapter{The Lantern That Remembered}

\section*{Told by}
Pilot Seref, a woman with one eye and a brass hook for a hand. She captains a little cutter called the *Red Wake*. Says the sea tells her where to go.

\section*{The Tale}

A young pilot lit a beacon on the headland. It was his first night alone. He trimmed the lamp, checked the mirror, poured oil. The flame burned steady.

At midnight, the beacon spoke. "I have seen this coast for three hundred years," it said. "I remember every ship that broke."

The pilot asked, "What do you want?"

"I want someone to listen. Then I want someone to change the light."

The pilot listened. He heard of wrecks, of storms, of pilots who died with their codes still secret. Then he climbed the tower and turned the lantern one degree east.

The beacon fell silent. The next morning, a ship that would have wrecked passed safely.

The pilot was promoted. He never told anyone about the voice. But every time he lit a new beacon, he spoke to it first. "I am listening," he said. "Tell me when to move."

\section*{The Witch's Closing}
"Beacons are not marks. They are arguments. If you do not listen, they will argue without you."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Speaking Beacon}

An old lighthouse whose flame is alive. It remembers every ship that has wrecked nearby. It can see storms days before they arrive. It will share its knowledge only with those who listen -- and who are willing to adjust the lantern's angle by a few degrees, changing the safe channel.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [KNOWLEDGE] [LOCATION]
\item A character who spends a night in the beacon and succeeds on Spirit + Lore (DV 4) gains a vision of the next storm's path (+1 die to navigation for the next leg).
\item The beacon may also reveal the location of a sunken wreck (a clue or treasure).
\item If the lantern is moved without the beacon's consent, the beacon goes dark for a month and the pilot gains a reputation as "deaf".
\end{itemize}

\paragraph{Adventure Hook:}
A coastal beacon has gone dark. Ships are wrecking on a reef that was never there before. The party must discover why the lighthouse is silent -- and whether the beacon is hiding something, or grieving something.

\section*{Colophon}
On the *Red Wake*, at anchor. The pilot spat into the waves. "She's been quiet since the war," she said. "I think she's waiting for someone."

