\chapter{The Bell That Never Rang}

\section*{Told by}
Bell‑warden Hollis, a gaunt man with soot under his fingernails and a cough that never quite stops. He tends the levee bells alone.

\section*{The Tale}

A bell was cast for a tower in the Mistlands. It was large, bronze, etched with ward‑runes. The smith said, “Ring it once, and the fog retreats. Ring it twice, and the dead listen. Ring it thrice, and the Direwood speaks.”

The warden hung it. He never rang it. He polished it. He oiled the clapper. He counted the hours without sound.

When the fog rose, the villagers pleaded. “Ring the bell!”

He refused. “It is not time.”

When a child was lost, they begged. “Ring it! The dead will help us find her.”

He refused. “The dead do not help. They collect.”

When the grey wraiths came, the villagers fled. The warden stayed. He climbed the tower. He did not ring the bell. He stood beside it, arms spread, and said nothing.

The wraiths passed. The tower stood. The bell never rang.

When they found the warden, he was dead. His hands were clasped around the clapper. He had kept it silent even as his breath left.

They buried him under the tower. The bell still hangs. No one has rung it.

\section*{The Witch's Closing}
“Some warnings should never be spoken. The bell knows. So did he.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Silent Bell}

A bell that rings only when its guardian dies. It will ring once, and its note is so pure that it banishes all fog and ghosts within a mile for a full day. After that, the bell cracks. It can be used only once.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [WARD] [SACRIFICE]
\item To ring the bell, a character must die while touching it. The bell tolls once, clearing all supernatural threats in a large area (GM discretion). The bell is then destroyed.
\item A character may attempt to ring it without dying by succeeding on a Body + Resolve test (DV 6) – but they suffer 3 Harm (and may still die). On success, the bell rings, they live, and the bell cracks but can be repaired (cost and time).
\end{itemize}

\paragraph{Adventure Hook:}
An old bell‑warden lies dying. The fog is rising, and the dead are walking. He has never told anyone the bell’s secret. The party must decide: ring it and lose it, or find another way – and perhaps become the next keepers of the silent tower.

\section*{Colophon}
Levee tower, morning. The warden coughed into his sleeve. “No,” he said, when I asked if he would ring it. “Not today.”

