\chapter{The Loom of Debts}

\section*{Told by}
Weaver Thalia, a woman with red‑stained fingers and a loom that never stops. She weaves shrouds for the guild.

\section*{The Tale}

A dyer bought a bolt of silk from a merchant. It was white, heavy, cool to the touch. The merchant said, "This silk remembers every shame it has witnessed. Weave with care."

The dyer did not believe him. He dyed the silk crimson and cut it into lengths for the market.

But the silk would not leave the loom. It rewound itself at night. The threads knotted into faces. The faces whispered confessions.

The dyer tried to burn it. The fire turned the silk black, and the confessional grew louder.

A witch came. She touched the loom. "You did not buy the silk," she said. "You bought the witness. It will not leave until you confess."

The dyer knelt. He admitted his shames -- the false weights, the watered dyes, the apprentice he had blamed for his own mistake.

The silk loosened. The witch cut a single thread. "This is your toll," she said. "Wear it as a bracelet. When you lie, it tightens."

The dyer wore the thread for the rest of his life. He never lied again. The silk was sold for a fair price -- and the buyer was told its nature.

\section*{The Witch's Closing}
"What you hide may hold you. Confess, or be woven into the cloth."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Witness Silk}

A bolt of silk that retains memory of every wrong witnessed in its presence. It cannot be destroyed by normal fire. To unbind it, someone must confess a secret shame in front of it. The confession frees one yard of silk at a time.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [TRUTH] [CURSE]
\item A character who possesses the silk gains +1 to Insight rolls (the silk whispers hints). But at each downtime, the GM may ask for a confession (a minor flaw, a forgotten debt). If refused, the silk tightens: the character suffers --1 die on all social rolls until they confess.
\item The silk can be "paid off" by sacrificing a yard (which becomes inert cloth) in exchange for a favor from a guild or spirit.
\end{itemize}

\paragraph{Adventure Hook:}
A merchant's best loom is stuck -- the silk will not come out. The merchant is hiding a secret that could ruin him. The party must decide: extract the confession, burn the loom, or find a third way.

\section*{Colophon}
Weaver's loft, dusk. The loom clicked. Thalia touched a thread. "This one is mine," she said. "I overcharged a widow. The silk remembers."

