\chapter{The Coin That Came Back}

\section*{Told by}
Old Salt, a Zakov pirate with a peg leg and a parrot that swears in three languages. He’s been drowned twice. The sea spit him out both times.

\section*{The Tale}

A smuggler found a coin in a dead man’s eye. It was silver, warm to the touch, stamped with a ship that had no name. He took it.

That night, the coin bought him safe passage past a patrol. The watchman looked at it and waved him through without a word.

The next day, the dead man’s ghost sat in his cargo hold. “You took my toll,” the ghost said. “Now you owe me passage.”

The smuggler tried to sell the coin. No one would take it. He tried to throw it overboard. It floated back. He tried to give it to a beggar. The beggar returned it at dawn.

He sailed to a shrine of the drowned. He placed the coin on the altar. The ghost appeared again. “Not enough,” it said. “You must carry me to the place I died.”

The smuggler sailed to a reef. He dropped the coin into the water. The ghost smiled, took the coin, and vanished.

The smuggler sailed home. His hold was full of fish – more than he had ever caught. He never picked up a coin from a dead man’s eye again.

\section*{The Witch's Closing}
“Not all treasure is worth its weight. Some coins carry a ticket. The ticket is always for where you do not want to go.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Ferry Coin}

A silver coin taken from a corpse’s eye. It cannot be spent, given away, or destroyed. It forces the holder to carry the ghost of the dead person to the place of their death. Once there, the coin and ghost vanish, and the holder is granted a boon of the sea (good fishing, calm waters, a safe passage).

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [SACRIFICE]
\item The holder gains +1 die on all Negotiation rolls (the coin whispers) but suffers nightmares every night unless they are actively sailing toward the ghost’s destination.
\item The ghost can be negotiated with – it might accept a substitute voyage (e.g., delivering a message instead of carrying the body). This requires a Will + Lore test (DV 5).
\end{itemize}

\paragraph{Adventure Hook:}
A sailor has a coin that won’t let him sleep. He doesn’t know whose ghost it is. The party must help him discover the dead person’s identity and their destination – before the ghost drives him mad.

\section*{Colophon}
Zakov tavern, the *Drowned Rat*. Old Salt flipped a coin to the barkeep. “That’s a live one,” he said. “Keep the change.”

