\chapter{The Veil and the Cradle}

\section*{Told by}
Vizier's widow, a woman who wears a mask even inside her own home. She speaks through a silk curtain. No one has seen her face in twenty years.

\section*{The Tale}

A slave girl was bought for the sultan's harem. She was young, frightened, and silent. The sultan's wife took pity on her.

"Wear a veil," the wife said. "The sultan will not see you. He sees only the silk. You may become invisible."

The girl wore the veil. She served the sultan's table. He did not see her. She cleaned his chambers. He did not remember her.

One night, the sultan's enemy sent a knife. The girl saw it. She could not call out -- her voice would break the veil. She stepped between the knife and the sultan.

She took the blade in her chest. The veil fell. The sultan saw her for the first time.

"Who are you?" he asked.

"The one who was never there," she said. Then she died.

The sultan had the veil woven into a shroud for her. He kept the shroud for the rest of his life. He never married again.

\section*{The Witch's Closing}
"The veil hides. The knife reveals. Choose which you carry."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Silent Veil}

A silk veil that makes the wearer unnoticeable -- not invisible, but forgettable. Guards will look past her; servants will not remember her face. She cannot speak while wearing it, or the effect breaks.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [STEALTH] [CURSE]
\item While wearing the veil, the character gains +2 dice to Stealth and Deception to avoid notice. However, they cannot speak (verbal spells, commands, conversation). Removing the veil is an action.
\item Each scene, the GM may ask the character to test Spirit + Resolve (DV 4). On failure, the character feels compelled to speak (loudly), breaking the veil and alerting nearby creatures.
\end{itemize}

\paragraph{Adventure Hook:}
A noble's daughter has been sold into a hostile court. She has a veil that makes her unnoticeable -- but she cannot speak. The party must smuggle her out without being seen, and without her breaking the silence.

\section*{Colophon}
A room behind a curtain. The widow poured tea. The cup rattled against the saucer. "I wore it for twenty years," she said. "Now I do not need it. The silence is inside me."

