\chapter{The Ford That Moved}

\section*{Told by}
Ferryman Dav, a Dhaharan with a turban and a voice like gravel. He poled his boat across the same ford for fifty years. Says the river changed course three times.

\section*{The Tale}

A caravan master needed to cross a ford. The river was high, the stones slippery. He prayed to the ford, offering a grain of wheat.

“I will stay still for one crossing,” the ford whispered. “Then you will owe me a name.”

He crossed. The ford held.

The next night, he returned. The river was different – deeper, swifter. “Name a price,” he said.

“You have already named it,” the ford replied. “The name of your firstborn child.”

The merchant had no child. He was childless. He began to weep.

The ford softened. “Then give me the name you would have given them. And every year, return and speak it.”

The merchant knelt and whispered a name – a name no one had ever heard. The ford calmed. He crossed safely.

He returned every year and spoke the name. His wife gave birth to a daughter. He named her the same name. The ford never troubled him again.

\section*{The Witch's Closing}
“A road is a covenant. Pay the toll honestly. The river remembers.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Naming Ford}

A river crossing that asks for a name as toll. It will accept a false name once – but the next crossing demands the true name of the traveller. If the traveller refuses, the ford shifts its course and becomes impassable for a year.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [THRESHOLD] [TRUTH]
\item To cross, a character must succeed on a Spirit + Lore test (DV 3) to recall a name that has meaning. On failure, the ford demands a true name of one of the party. That character then gains a String on themselves (the ford knows their name) and can be compelled by water magic.
\item If the party tries to cross without paying, they must navigate the ford as a hazard (Athletics DV 5) – on failure, they are swept downstream and lose 1 Supply.
\end{itemize}

\paragraph{Adventure Hook:}
A ford has closed to all trade. The local lord sent a surveyor; the surveyor returned, but he cannot remember his own name. The party must learn what the ford took – and whether they can pay it back.

\section*{Colophon}
Ferryman’s hut, near the saffron fields. Dav spat into the water. “She’s quiet tonight,” he said. “Must have taken a good one.”

