\chapter{The Fog Road}

\section*{Told by}
Mist‑walker Kell, an old Ubral woman who lives in a cairn and never lights a fire. Says the fog is not a thing; it is a question.

\section*{The Tale}

A child got lost on the moor. The fog came down thick and grey. She could not see her feet. She walked for hours, but always returned to the same stone.

A voice spoke from the mist. "Who walks on my road?"

"I do," said the child. "I am lost."

"Are you?" said the voice. "You are walking exactly where I need you. But you have not answered my question. Who walks?"

The child thought. She could not give her name -- names are for daytime. She could not say her clan -- clans are for fires.

"Someone who belongs to the fog," she said.

The voice laughed. "Then you may stay."

The fog parted. The child saw a path of white stones. She followed it home. At the edge of the moor, she looked back. The fog was gone.

Her grandmother said, "You walked with the Urisk. It asks every traveller the same question. Most answer with their names. Then it keeps them."

\section*{The Witch's Closing}
"The fog does not hide. It asks if you belong. If you do, it will guide you. If you do not, it will keep you."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Urisk of the Moor}

A spirit of the fog. It appears to lost travellers and asks, "Who walks on my road?" If the traveller gives a name, the Urisk learns it and may use it to compel the traveller to stay. If the traveller claims belonging to the fog, the Urisk guides them safely and may become an ally.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [SPIRIT] [THRESHOLD]
\item To negotiate with the Urisk, a character must succeed on a Spirit + Lore test (DV 4) while offering a truth about themselves that is not their name (a memory, a regret, a hope). On success, the Urisk parts the fog and the character may re‑roll any failed navigation check on the moor.
\item If the character gives their name, the Urisk can later appear as a difficult choice: a Path that seems easy but leads to danger, etc.
\end{itemize}

\paragraph{Adventure Hook:}
A shepherd's child has vanished into the fog. The villagers are afraid to search. The party must find the Urisk and answer its question -- without losing themselves.

\section*{Colophon}
Cairn on the high moor. The old woman did not light a candle. "Light attracts the wrong things," she said. "The dark is easier to negotiate with."

