\chapter{The Counting of the Fifth Breath}

\section*{Told by}
Sergeant Relled, a Black Banner mercenary with a missing ear and a habit of counting his steps. He has survived seven campaigns. He does not know why.

\section*{The Tale}

A young captain lost his Blue in a skirmish. The enemy had five breaths to plant a banner. The captain panicked.

“I have four breaths left,” he told his sergeant. “What do I do?”

The sergeant pointed at a ridge. “How many ways up?”

“Two.”

“Then they are already dead. Weld one. Block the other. Use the first breath to drink water.”

The captain drank. The sergeant welded the path with a line of reds. The enemy runner hesitated. The captain’s second breath, he moved an orange to fork the ridge. Third breath, he planted a green on the low path. Fourth breath, he waited.

The enemy runner never reached the ridge. The fifth breath expired. The captain won.

The sergeant said, “Panic is a debt you cannot afford. Count the breaths. Then count your exits. Then count your water. Then move.”

\section*{The Witch's Closing}
“Re‑forge is a budget, not a panic. Spend your first breath on calm. The rest will follow.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Five‑Breath Clock}

A magical timer that counts down from five (minutes, turns, actions). It appears when a character is in a desperate race – a siege, a chase, a ritual. The character can see the count (a floating number). Each time they act decisively, the clock pauses for a heartbeat. Hesitation makes it tick faster.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [PRESSURE] [DREAD]
\item When the clock appears, the character gains +1 die to all actions related to the objective (panic focus) but suffers –1 die to all other actions (tunnel vision).
\item If the clock reaches zero, the objective fails. The character can reset the clock by finding a new approach (e.g., a side path, a different tactic), which requires a Vision + Tactics test (DV 4).
\end{itemize}

\paragraph{Adventure Hook:}
A mercenary company is trapped in a collapsing fortress. They have five minutes before the enemy arrives. The captain has a magical hourglass that ticks audibly. The party must complete their escape – or find a way to break the clock.

\section*{Colophon}
Campfire near the Singing Wastes. The sergeant counted his fingers. “Five,” he said. “Always check after a fight. You never know when one went missing.”

