\chapter{The Saffron Gate}

\section*{Told by}
Merchant Zara, a Fhara woman who wears bells on her ankles and never haggles at the first offer. She says patience is a type of geometry.

\section*{The Tale}

A merchant came to the Saffron Gate. It was high, carved with lions, guarded by two men who did not blink.

"Toll is one truth," said the gate.

The merchant offered, "I am honest."

The gate did not open.

He offered, "My goods are pure."

The gate did not open.

He offered, "I have never cheated a customer."

The gate swung open. Inside, a woman sat on a stool. "You have just told three truths," she said. "The first was a boast. The second was a claim. The third was a lie. The gate only opens for lies."

The merchant stared. "Then what is the toll?"

"The knowledge that you are a liar," said the woman. "Now pass. And remember: the gate knows when you believe your own stories."

\section*{The Witch's Closing}
"A lie that you believe is not a lie. It is a cage. The gate opens from the inside."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Truth Gate}

A magical gate that opens only when a character speaks a falsehood they genuinely believe -- not a deliberate lie, but a self‑deception. Once opened, the character gains a small epiphany (a realisation that they were wrong about something important). The gate can be used only once per person.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [TRUTH] [THRESHOLD]
\item A character must declare a "truth" about themselves that is actually false. The GM may ask the player to describe the belief. If the GM judges that the character genuinely believes it, the gate opens. The character then gains +1 die to Insight rolls regarding that specific subject (they have seen their own blind spot).
\item The gate will not open for obvious lies or boasts.
\end{itemize}

\paragraph{Adventure Hook:}
A diplomat has been sent to open a sealed pass -- but the gate requires a self‑deception. The diplomat believes she is the best negotiator in the land. She is not. The party must help her see the truth, or find a way to trick the gate.

\section*{Colophon}
Saffron bazaar, afternoon. Zara counted coins into eight piles. The ninth she left empty. "For the lie I believe," she said.

