\chapter{The Grain That Spoke}

\section*{Told by}
Granary‑keeper Oris, a heavy man with a gentle voice and a ledger that never balances. Says the silo talks to him when the wheat is low.

\section*{The Tale}

A clerk was tallying grain in the silo. He counted aloud: one sack, two sacks, three, four, five, six, seven, eight.

“Ninth,” whispered the grain.

He stopped. He had been taught: never count the ninth. He looked. The grain had shifted. A single kernel lay on the floor.

He picked it up. It was warm. He put it in his pocket.

That night, he dreamed of debtors. Their names were scratched into his ledger. He woke and forgave three.

The next day, the grain was heavier. He counted again. Eight sacks. The ninth was full of light.

He forgave another debtor. The grain remained full.

For a year, he forgave one debtor each night. The granary never emptied. The villagers called it a miracle. The clerk said, “It is a conversation.”

When he died, they found the kernel in his hand. It had grown into a small, green shoot. They planted it by the silo. The grain never spoke again.

\section*{The Witch's Closing}
“Mercy enriches the giver. The grain knows. It counts only the eighth name – the one you chose to spare.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Speaking Grain}

A silo of grain that whispers the names of those who are in debt. The grain can be used to “pay” a debt – by forgiving the debtor, the grain replenishes itself. If the owner uses the grain for profit instead of mercy, the silo empties.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [HEAL] [TRUTH]
\item Once per week, the grain reveals one name of someone who owes a debt (material, emotional, or magical). The owner may choose to forgive that debt. The grain then produces 1 Supply (as if a new shipment arrived).
\item If the owner chooses to collect the debt instead, the grain becomes stale and provides no benefit for a month.
\end{itemize}

\paragraph{Adventure Hook:}
A village granary is overflowing despite a drought. The keeper is growing rich – but he has stopped forgiving debts. The party must learn the secret of the grain and decide whether the miracle is worth the price.

\section*{Colophon}
Granary, evening. Oris held up a single kernel. “This is the first one,” he said. “It still speaks. But I do not listen anymore.”

