\chapter{The Bargain of the Unlit Lamp}

\section*{Told by}
A Trow who gave no name. It sat in the corner of the tavern, smoking a pipe that had no ember. Its voice was dry leaves. I wrote what I heard.

\section*{The Tale}

A traveller came to a crossroads at midnight. A figure waited there. It wore a grey robe and a hat that drank the starlight.

“I am a Trow,” it said. “I trade in absence. What do you lack?”

The traveller lacked nothing. He was wealthy, happy, loved.

“Then sell me something you do not need,” said the Trow. “A memory. A breath. A name you have never used.”

The traveller laughed and sold his left‑handed glove.

The Trow gave him a lamp that would never go out. “The light will not fade,” it said. “But you will forget why you needed it.”

The traveller took the lamp. He walked home. The light guided him. But when he arrived, he could not remember his wife’s face. He could not remember the way to his own door.

He returned to the crossroads. The Trow was gone. He found his glove on the ground. He put it on. The lamp went out.

He sat in the dark, remembering. The next morning, he left the lamp at the crossroad and walked away. He never bought from a Trow again.

\section*{The Witch's Closing}
“Trow trade in gaps. The gap that light fills is the gap where memory used to be.”

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Trow Lamp}

A lantern that never needs oil and never goes out – but each hour it burns, the user loses a random memory (a face, a word, a route). Once lit, it cannot be extinguished except by returning the price paid (which the Trow will accept only once).

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [LIGHT] [CURSE]
\item The lamp provides perfect illumination in a 30‑foot radius. Every hour, the GM erases one minor fact from the character’s background (the character may still remember, but the fact is “blank” – they cannot recall it under pressure). After 8 hours, the character cannot remember their own name.
\item To extinguish the lamp, the character must give up something of equal value to what they originally paid (e.g., a cherished possession, a close relationship, a skill point).
\end{itemize}

\paragraph{Adventure Hook:}
A lighthouse keeper uses a lamp that never dies. But the keeper has forgotten his own daughter’s name – and the sea is rising. The party must convince him to extinguish the light, or find a way to pay the Trow’s price.

\section*{Colophon}
Tavern corner, after midnight. The Trow smiled. Its teeth were grey. “I will take that question,” it said. “You have already forgotten you asked it.” I looked at my notes. The page was blank.

