\chapter{The Soup That Was Free}

\section*{Told by}
Hag Morag (or so she called herself). She sat by a well, stirring a pot that had no fire under it. The soup was hot. She did not explain.

\section*{The Tale}

A starving family came to a cottage. A crone answered the door. "I have soup," she said. "It is free."

They ate. The soup was rich, meaty, warm. They ate until they were full.

The crone watched. "The soup is free," she said again. "But the seconds will cost you."

The family hesitated. The father said, "We have no coin."

"I do not need coin," said the crone. "I need a breath. One breath from each of you. Hold it for me. I will give it back when you next share a meal."

They agreed. They breathed into her kettle. The kettle steamed.

They left. That night, they dreamed they were drowning. They woke gasping. Their breath was shorter.

They returned to the cottage. The crone was stirring the same pot. "The debt is not due yet," she said. "Eat again. The soup is still free."

They ate. They were not hungry. They ate anyway.

The crone took another breath. The kettle grew heavier.

For a year, they ate free soup once a week. Each time, the crone took a breath. Each time, their sleep grew deeper, their waking shorter.

On the last night, the father said, "We cannot pay. We have no breath left."

The crone nodded. "Good," she said. "Now you are honest. The debt is paid. The soup was the toll. The breaths were the interest. Go home. Cook your own meals."

They went home. They felt lighter. Their breath returned.

They never ate free soup again.

\section*{The Witch's Closing}
"Free soup is a hook. The hook is not the soup. It is the second bowl."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Hungry Kettle}

A pot that produces unlimited nourishing soup -- but each bowl consumed also takes a small amount of breath (Fatigue). After enough bowls, the eater loses the ability to regain Fatigue naturally (they require magical healing or a specific service to the Hag).

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [HEAL] [CURSE]
\item The soup restores 1 Supply per bowl. Each bowl also causes the eater to mark 1 Fatigue (which cannot be recovered by rest until the curse is lifted).
\item Clock: \textit{Hag's Debt} [4] -- each bowl ticks once. When full, the eater must perform a favor for the Hag (or a hag of her choosing). Refusal causes the Fatigue to become permanent (cannot be healed).
\end{itemize}

\paragraph{Adventure Hook:}
A village has been living on free soup from a mysterious old woman. The villagers are growing weak, sleeping too much. The party must discover the price of the soup -- and whether they can break the kettle without starving the village.

\section*{Colophon}
Well at a crossroad. The crone winked. "You did not eat," she said. "That is good. I would have taken your ink."

