\section{Oath of Flame and Light}

\subsection{Domain and Epithets}

The Oath of Flame and Light is no patron of half-measures. Its fire names, binds, and burns—demanding that those who swear within its radiance stand openly, speak truly, and pay the cost of keeping their word. At dawn altars, the sworn kindle sparks of consecrated fire; in battle, they blaze as torches that hold back the night. To follow this Oath is to live in public truth, with no shadow to hide in and no retreat from the vow once spoken.

\subsection{Cultural Origins}

The Oath appears in many guises across the known world:

\begin{itemize}
\item In \textbf{Aeler}, it is the Everflame, carried from shrine to shrine by wandering priests. Its censers never cool.
\item In \textbf{Vhasia}, it is Adur, the Light of Aeler, whose flame consecrates oaths and consumes falsehood. Dwarven templars strike their vows into stone with a brand.
\item In \textbf{Viterra}, it is the Dawnfire, invoked as a war-god of vows and vengeance, whose rising light marks those who break faith. Cavalrymen light a single arrow before battle.
\item In \textbf{Ecktoria}, it is the Watchfire, a saint's flame that never gutters, tended by a silent order of healers who swear to harm no one—and to let no one die alone.
\end{itemize}

Wherever it manifests, the Oath remains the same: to swear beneath its dawn is to bind oneself to a truth that cannot be hidden, forgotten, or undone.

\subsection{Warrior Archetype: The Dawn Warrior}

A Runic Warrior of the Oath of Flame and Light is not a scholar or a courtier. They are a \textit{Dawn Warrior}—a crusader who carries their vows in steel and fire. Their Codex is engraved on a gorget or gauntlet. Their Thiasos is a warhorse with eyes like embers or a spirit-hound of ash and seal-wax. Their Rites are not whispered; they are declared.

\subsection{Codex Form}

The Dawn Warrior's Codex is a physical object worn into battle:

\begin{itemize}
\item \textbf{Gorget:} A steel collar engraved with vows. Worn over the throat, protecting the voice that speaks the oath.
\item \textbf{Gauntlet:} A plate glove marked with runes of dawn. The hand that strikes also swears.
\item \textbf{Branded Bracer:} A leather or steel armguard, each brand a Rite invoked and paid.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Warhorse:} A horse with eyes like dying embers. It fears no flame and will not bolt from battle.
\item \textbf{Spirit-Hound of Ash:} A hound whose fur smolders and whose paws leave no tracks. It can sense oath-breakers within Near range.
\item \textbf{Dawnblade (Sentient):} A sword that hums with trapped light. It speaks in proverbs about honor. It refuses to strike an unarmed foe.
\end{itemize}

\subsection{Core Mechanic: The Witnessed Vow}

When a Dawn Warrior swears an oath before a witness, they gain +1 die to all actions that serve that oath until it is fulfilled or broken. This stacks with the Unbroken Vow talent.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Kindled Blade (4 XP)}

\textbf{Tags:} [FIRE], [STRIKE], [WITNESS]

\textbf{Threshold:} Dawn, or a flame lit at dawn.

Your weapon ignites with radiant light. For one scene, your melee attacks deal +1 Harm against undead, oath-breakers, and creatures of corruption. The flame does not spread, but it illuminates a Close radius.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Steadfast Gaze (4 XP)}

\textbf{Tags:} [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A flame between you and the target.

Your eyes catch the light. For one exchange, a single target within Near range cannot lie to you. They may refuse to answer, but any statement they make must be true.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Unbroken Vow (7 XP)}

\textbf{Tags:} [VOW], [WARD], [WITNESS]

\textbf{Threshold:} A flame lit at dawn, and a witness who is not the Runekeeper.

Swear an oath before the flame and the witness. For the duration of the oath, you gain +1 Armor. Any ally who hears the oath and aids you gains +1 die to actions that serve it.

If you break the oath, you immediately mark 2 Harm (Stress) and the flame gutters. The witness may release you from the oath, but only by speaking your name aloud.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of Dawn's Judgment (8 XP)}

\textbf{Tags:} [FIRE], [STRIKE], [JUDGMENT], [WITNESS]

\textbf{Threshold:} A flame lit at dawn, and a target who has broken an oath.

Your blade blazes with dawn fire. Make a single melee attack. On a hit, the attack ignores 1 point of Armor and deals +2 Harm to undead, oath-breakers, and creatures of corruption. On a miss, the flame gutters but does not go out.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Dawnfire Covenant (12 XP)}

\textbf{Tags:} [VOW], [AREA], [WARD], [WITNESS]

\textbf{Threshold:} A flame lit at dawn, and at least three witnesses.

Consecrate a zone (Near radius) as dawnfire ground. For the scene:
\begin{itemize}
\item Allies within the zone gain +1 die to resist fear.
\item Enemies who are undead, oath-breakers, or creatures of corruption suffer -1 die to all actions while within the zone.
\item Any oath sworn within the zone is witnessed by the flame itself. Breaking it causes the oath-breaker to mark 1 Harm (Stress) immediately, regardless of distance.
\end{itemize}

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Dawn's Judgment (6 XP)}

\textbf{Requirement:} Oath of Flame and Light, Body 3+

When you invoke Rite of Dawn's Judgment, you may choose to extend the effect. Your blade remains kindled for the remainder of the scene, not just a single strike. Each additional attack beyond the first costs 1 Fatigue.

\subsubsection{Unquenchable (4 XP)}

\textbf{Requirement:} Oath of Flame and Light

Your flame cannot be extinguished by mundane means. Wind, water, and darkness do not affect your kindled blade or dawnfire Rites. Magical extinguishing requires a contested roll (Spirit + Lore vs. your Spirit + Presence).

\subsubsection{The Light Reveals (8 XP)}

\textbf{Requirement:} Oath of Flame and Light, Witnessed Scar

Your scars are visible even through armor—glowing faintly where dawnfire has marked you. Once per session, you may touch a scar and ask the GM one question about an oath-breaker or creature of corruption within Far range. The GM must answer truthfully, though the answer may be cryptic.

\subsection{Roleplaying a Dawn Warrior}

\begin{itemize}
\item You do not whisper your vows. You declare them.
\item You do not hide your covenant. You wear it.
\item You may struggle with mercy. The flame judges; you must learn when to temper it.
\item You may be called to witness oaths outside of battle—marriages, treaties, adoptions. Your word carries weight.
\end{itemize}

\subsection{Sample NPC: Knight-Commander Elara of the Dawn Vigil}

\textbf{Archetype:} Dawn Warrior

\textbf{Patron:} Oath of Flame and Light

\textbf{Codex:} Engraved gorget, each vow a line of flame-light script

\textbf{Thiasos:} Ember, a spirit-hound of ash whose paw prints smolder for hours

\textbf{Personality:} Stern, weary, secretly uncertain. She has kept every vow she ever swore. She is not sure she remembers why.

\textbf{Conflict:} A vow she swore twenty years ago—to protect a village that no longer exists—has no witness but herself. Does the oath still bind? The flame says yes. Her heart is not sure.

\subsection{GM Guidance: Intrusions for Oath of Flame and Light}

\begin{itemize}
\item A vow the Dawn Warrior swore years ago resurfaces. A witness they had forgotten appears.
\item Their flame gutters at a crucial moment—not because of magic, but because of doubt.
\item An enemy begs for mercy. The flame demands judgment. The Dawn Warrior must choose.
\item A dawnfire Rite leaves a permanent mark: a scar that glows faintly in darkness, visible to all who know what it means.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
