\section{Rites of the Vowed}

The Vowed draw their power from oaths, law, and sacred judgment. Their Rites are declarations as much as invocations—spoken aloud, witnessed, and binding. These Rites are available to Runic Warriors of the Oath of Flame and Light, the Oath of Mercy and Grace, Mykkiel, the Inquisitor Prime, and the Witness.

\subsection{Oath of Flame and Light — Dawn Warrior}

\subsubsection{Low Rite: Rite of the Kindled Blade (4 XP)}

\textbf{Tags:} [FIRE], [STRIKE], [WITNESS]

\textbf{Threshold:} Dawn, or a flame lit at dawn.

Your weapon ignites with radiant light. For one scene, your melee attacks deal +1 Harm against undead, oath-breakers, and creatures of corruption. The flame does not spread, but it illuminates a Close radius.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Steadfast Gaze (4 XP)}

\textbf{Tags:} [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A flame between you and the target.

Your eyes catch the light. For one exchange, a single target within Near range cannot lie to you. They may refuse to answer, but any statement they make must be true.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Unbroken Vow (7 XP)}

\textbf{Tags:} [VOW], [WARD], [WITNESS]

\textbf{Threshold:} A flame lit at dawn, and a witness who is not the Runekeeper.

Swear an oath before the flame and the witness. For the duration of the oath, you gain +1 Armor. Any ally who hears the oath and aids you gains +1 die to actions that serve it.

If you break the oath, you immediately mark 2 Harm (Stress) and the flame gutters. The witness may release you from the oath, but only by speaking your name aloud.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of Dawn's Judgment (8 XP)}

\textbf{Tags:} [FIRE], [STRIKE], [JUDGMENT], [WITNESS]

\textbf{Threshold:} A flame lit at dawn, and a target who has broken an oath.

Your blade blazes with dawn fire. Make a single melee attack. On a hit, the attack ignores 1 point of Armor and deals +2 Harm to undead, oath-breakers, and creatures of corruption. On a miss, the flame gutters but does not go out.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Dawnfire Covenant (12 XP)}

\textbf{Tags:} [VOW], [AREA], [WARD], [WITNESS]

\textbf{Threshold:} A flame lit at dawn, and at least three witnesses.

Consecrate a zone (Near radius) as dawnfire ground. For the scene: allies within the zone gain +1 die to resist fear; enemies who are undead, oath-breakers, or creatures of corruption suffer -1 die to all actions while within the zone; any oath sworn within the zone is witnessed by the flame itself.

\textbf{Obligation:} 3

\subsection{Oath of Mercy and Grace — Hospitaller}

\subsubsection{Low Rite: Rite of the Shared Bowl (4 XP)}

\textbf{Tags:} [HEAL], [BOND], [RESTORATION]

\textbf{Threshold:} A bowl of clean water, shared with a willing creature.

You and the creature each clear 1 Fatigue. The creature gains +1 die on their next roll to resist fear or despair.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Grey Touch (4 XP)}

\textbf{Tags:} [HEAL], [CLEANSE]

\textbf{Threshold:} A wound that has drawn blood.

Touch a wounded creature. Remove 1 Fatigue or stabilize a dying patient (they will not die from this wound for the rest of the scene). The wound closes slightly; the pain eases.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Sheltered Hearth (7 XP)}

\textbf{Tags:} [WARD], [AREA], [SANCTUARY]

\textbf{Threshold:} A door, a threshold, or a hearth.

Consecrate a zone (Near radius) as sanctuary. For the scene: no violence may be initiated within the zone; the dying cannot die while within the zone. If the Rite is broken (by an act of violence from within), the Runekeeper marks 2 Fatigue.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Spiral's Witness (8 XP)}

\textbf{Tags:} [DEBT], [RESTORATION], [FEAR], [TRUTH]

\textbf{Threshold:} A grey wool cord tied around the Runekeeper's wrist.

Name a person crushed by debt or guilt. For one scene, they ignore all penalties from unpaid obligations. They are not free; they are unburdened. If that person then commits an act of violence against an innocent, the unburdening reverses and they suffer double penalties.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Endless Spiral (12 XP)}

\textbf{Tags:} [HEAL], [RESTORATION], [DEBT], [AREA], [FEAR]

\textbf{Threshold:} A circle of nine grey stones, a bowl of tears, and a candle lit at each cardinal point.

All within the circle who accept the Oath's blessing clear 2 Fatigue and remove one Condition caused by despair or guilt. For the rest of the scene, any act of care performed within the zone costs the giver -1 Obligation (minimum 0). Any creature that harms a caregiver within the zone gains a Promise Clock [6] named "Spiral's Debt."

\textbf{Obligation:} 3

\subsection{Mykkiel — Justicar}

\subsubsection{Low Rite: Rite of the Measured Debt (4 XP)}

\textbf{Tags:} [VOW], [COMMAND], [WITNESS]

\textbf{Threshold:} A pair of scales balanced with tokens from both sides.

Establish a temporary accord. Both parties suffer -1 die if they break it first. You gain +1 die to enforce compliance.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Weighed Heart (5 XP)}

\textbf{Tags:} [TRUTH], [SENSE], [WITNESS]

\textbf{Threshold:} A small brass scale touched briefly to the chest.

Sense if the target acts against their nature or oath. Gain +1 die when pressing them. If the target is lying, the scale tips—only you can see it.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Balanced Scales (8 XP)}

\textbf{Tags:} [DEBT], [BOND], [TRANSFER], [WITNESS]

\textbf{Threshold:} A set of scales inscribed with runes of parity.

Exchange a burden between two willing parties (Harm for Fatigue, Debt for Favor, etc.). Both gain +1 die to cooperate. If the exchange is unwilling, you must win a contested Command roll (Presence + Command vs. Resolve).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Covenant Seal (9 XP)}

\textbf{Tags:} [VOW], [BOND], [DEBT], [WITNESS]

\textbf{Threshold:} A written contract and a flame to seal the wax.

Formalize a binding covenant between two or more parties. Each party must sign or mark the document. Violation of the covenant causes the violator to mark 1 Harm (Stress) immediately, and the GM gains 2 SB.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Final Adjudication (13 XP)}

\textbf{Tags:} [AREA], [COMMAND], [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A circle of iron bells, each etched with runes of judgment.

Convene an unseen tribunal. For the next scene, disputes within the zone are judged formally: +2 dice to Command when speaking as arbiter, and honest testimony gains +1 die. Any creature that lies within the zone suffers 1 Harm (Stress).

\textbf{Obligation:} 3

\subsection{Inquisitor Prime — Purifier}

\subsubsection{Low Rite: Rite of the Pure Flame (4 XP)}

\textbf{Tags:} [PROTECT], [RADIANT], [FEAR], [WITNESS]

\textbf{Threshold:} Ash from a burned grimoire or a sigil-scarred candle.

Gain +2 dice to resist supernatural influence (fear, charm, possession) and +1 Armor vs. magical attacks. Undead, demonic, or corrupted foes suffer -1 die against you.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Unclouded Eye (5 XP)}

\textbf{Tags:} [TRUTH], [KNOWLEDGE], [RADIANT], [WITNESS]

\textbf{Threshold:} Silver dust stirred into holy water.

Gain +2 dice to Investigation or Insight when spotting illusions, glamours, or magical deception. You cannot be surprised by sorcery this scene.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Cleansing Light (8 XP)}

\textbf{Tags:} [DISPEL], [RADIANT], [FORCE], [WITNESS]

\textbf{Threshold:} A shard of blessed mirror and consecrated oil.

Target a magical effect (ward, curse, enchantment). Test Spirit + Lore vs. DV = effect's Tier. On a hit, the effect is dispelled. On a partial, it is weakened. On a miss, the effect lashes back—GM gains 1 SB.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Marked Prey (7 XP)}

\textbf{Tags:} [CURSE], [TRACK], [FEAR], [WITNESS]

\textbf{Threshold:} A belonging of the target mixed with blessed salt.

Mark a supernatural target: they suffer -1 die to all rolls and cannot hide with illusion. You gain +1 die to track or strike them. Undead, demonic, or corrupted foes take +1 Harm from you.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Final Admonition (13 XP)}

\textbf{Tags:} [DEATH], [BANISH], [FORCE], [FEAR], [WITNESS]

\textbf{Threshold:} Ash from a destroyed spellbook, martyr's water, and the target's true name.

Attempt to annihilate one supernatural foe. Target tests Spirit + Resolve (DV 5). On a failure, they are destroyed outright. On a success, they suffer Harm 3 and -2 dice this scene. This Rite may be used only once per target.

\textbf{Obligation:} 3

\subsection{The Witness — Truth-Sworn}

\subsubsection{Low Rite: Rite of the Lingering Trace (4 XP)}

\textbf{Tags:} [INVESTIGATION], [KNOWLEDGE], [TRUTH]

\textbf{Threshold:} A trace of the thing to be remembered (hair, dust, a spoken name).

Gain +1 die to investigate or notice something directly related to the trace within the current scene. Ask the GM one yes/no question about whether the trace is connected to a specific person, event, or object.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Uncomfortable Question (5 XP)}

\textbf{Tags:} [COMMAND], [FEAR], [TRUTH], [WITNESS]

\textbf{Threshold:} A circle drawn with chalk or string while focusing on a specific person.

Ask one question of a target within the zone. They may either answer truthfully (the GM must give a true answer, though it may be cryptic) or mark 1 Fatigue and gain a \textit{Guilty Tell} Condition.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Echoing Truth (8 XP)}

\textbf{Tags:} [MIND], [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A reflective surface (mirror, still water, polished metal) used to focus on the target.

The target tests Resolve (DV 3) or suffers -1 die to rolls involving memory, deception, or resisting interrogation for the scene. If they fail, you may ask one specific, factual question about something they know, and they must answer truthfully or suffer 1 SB.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Immutable Record (7 XP)}

\textbf{Tags:} [VOW], [BOND], [TRUTH], [WITNESS]

\textbf{Threshold:} A document signed by all parties within the zone, or a spoken pact witnessed by the caster.

Bind the agreement. Any party who knowingly breaches it suffers 1 SB (Hearts) immediately and gains a persistent \textit{Oathbreaker's Mark} Condition.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Final Reckoning (13 XP)}

\textbf{Tags:} [AREA], [COMMAND], [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A formally called gathering (court, council, family meeting) within the consecrated zone.

All present must speak their greatest debt, wrongdoing, or hidden truth related to the gathering's purpose. Those who lie or withhold suffer Harm 2 (Stress/Reputation). Truth-tellers gain +2 dice to social actions for the remainder of the scene within the zone.

\textbf{Obligation:} 3

The thread held. The water carried. That is not a Rite. That is grace.
