\section{Oath of Mercy and Grace}

\subsection{Domain and Epithets}

The Oath of Mercy and Grace has no founder, no scripture, no single temple. It began in a hundred places—a battlefield where a medic bound wounds without asking whose side; a plague village where a stranger stayed to dig graves; a debtor's cell where a guard looked away and said, "I saw nothing." Those who swear the Oath do not worship a power. They embody a principle: that every act of care creates an obligation—not to the giver, but to the next person in need. You do not pay back. You pay forward. The mercy spirals.

But mercy is not weakness. The spiral does not forget. Those who strike a healer, who violate a sanctuary, who refuse the offered cup—they find that grace has teeth, and forgiveness carries a price.

\subsection{Cultural Origins}

The Oath appears wherever healers, midwives, and hospice workers have organized to protect their own:

\begin{itemize}
\item In \textbf{Viterra}, the Grey Spiral runs hospices along the border roads. They ask no payment, but they record every name in a ledger that never dries.
\item In \textbf{Aelaerem}, the Hearth-Menders tend the sick and settle disputes. Their symbol is a red thread tied around the wrist.
\item In \textbf{Sidhi} ports, the Weavers of Mercy treat wounded sailors and refugees. They are tolerated because they ask nothing—and because their hospices are the only neutral ground in a hundred miles.
\item In \textbf{Ecktoria}, the Order of the Salted Veil tends the dying in the shadow of the Everflame. The Church does not officially recognize them. The Church does not interfere.
\end{itemize}

\subsection{Warrior Archetype: The Hospitaller}

A Runic Warrior of the Oath of Mercy and Grace is not a crusader. They are a \textit{Hospitaller}—a warrior who fights not to kill, but to protect. Their Codex is a grey cloak or a shield marked with the spiral. Their Thiasos is a grey cord that tightens when mercy is refused, or a hospice bird that sings when someone is about to die. Their Rites bind wounds, create sanctuary, and turn an enemy's violence back upon itself.

\subsection{Codex Form}

The Hospitaller's Codex is a symbol of their calling:

\begin{itemize}
\item \textbf{Grey Cloak:} Woven from undyed wool, each fold a record of a mercy given. The cloak is heavy with memory.
\item \textbf{Shield of the Spiral:} A wooden or steel shield painted with a grey spiral. Rites are carved along the rim.
\item \textbf{Locket:} A silver locket containing a lock of hair from a patient who could not be saved. It is heavy. It reminds.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Grey Cord:} A length of undyed wool rope. It tightens when a promise is broken. It loosens when mercy is accepted. It can be tied around a wrist to witness an oath.
\item \textbf{Hospice Bird:} A small grey bird, often a dove or sparrow. It sings when someone within Near range is about to die. It does not sing for enemies or allies—only for the dying.
\item \textbf{The Silent Bell (Sentient):} A small bronze bell that does not ring—it hums. It speaks in whispers, and only to remind the Runekeeper that mercy is not weakness.
\end{itemize}

\subsection{Core Mechanic: The Spiral}

When a Hospitaller shows mercy to an enemy (sparing a life, binding a wound, offering sanctuary), they gain 1 Boon. This Boon can be used only for acts of protection, healing, or sanctuary.

When a Hospitaller is harmed by someone they have shown mercy, they may spend 1 Boon to force the attacker to test Resolve (DV 4) or suffer -1 die to all actions for the scene (the weight of their cruelty).

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Shared Bowl (4 XP)}

\textbf{Tags:} [HEAL], [BOND], [RESTORATION]

\textbf{Threshold:} A bowl of clean water, shared with a willing creature.

You and the creature each clear 1 Fatigue. The creature gains +1 die on their next roll to resist fear or despair.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Grey Touch (4 XP)}

\textbf{Tags:} [HEAL], [CLEANSE]

\textbf{Threshold:} A wound that has drawn blood.

Touch a wounded creature. Remove 1 Fatigue or stabilize a dying patient (they will not die from this wound for the rest of the scene). The wound closes slightly; the pain eases.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Sheltered Hearth (7 XP)}

\textbf{Tags:} [WARD], [AREA], [SANCTUARY]

\textbf{Threshold:} A door, a threshold, or a hearth.

Consecrate a zone (Near radius) as sanctuary. For the scene:
\begin{itemize}
\item No violence may be initiated within the zone. Existing violence continues, but new acts of aggression require a Resolve test (DV 4).
\item The dying cannot die while within the zone. They remain stable, suspended between breath and silence, until the Rite ends.
\end{itemize}

If the Rite is broken (by an act of violence from within), the Runekeeper marks 2 Fatigue.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Spiral's Witness (8 XP)}

\textbf{Tags:} [DEBT], [RESTORATION], [FEAR], [TRUTH]

\textbf{Threshold:} A grey wool cord tied around the Runekeeper's wrist.

Name a person crushed by debt or guilt. For one scene, they ignore all penalties from unpaid obligations (social, legal, or spiritual). They are not free; they are unburdened. If that person then commits an act of violence against an innocent, the unburdening reverses and they suffer double penalties.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Endless Spiral (12 XP)}

\textbf{Tags:} [HEAL], [RESTORATION], [DEBT], [AREA], [FEAR]

\textbf{Threshold:} A circle of nine grey stones, a bowl of tears, and a candle lit at each cardinal point.

All within the circle who accept the Oath's blessing clear 2 Fatigue and remove one Condition caused by despair or guilt. For the rest of the scene, any act of care performed within the zone costs the giver -1 Obligation (minimum 0). Additionally, any creature that harms a caregiver within the zone gains a Promise Clock [6] named "Spiral's Debt." When full, they suffer Harm 2 (Stress) as the weight of their cruelty crushes them.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Spiral Counter (6 XP)}

\textbf{Requirement:} Oath of Mercy and Grace, Body 3+

When a creature with an active Promise Clock (from any of your Rites) attacks you, you may spend 1 Boon to advance that clock by 1 segment immediately, before the attack resolves.

\subsubsection{The Unbuckled Gorget (4 XP)}

\textbf{Requirement:} Oath of Mercy and Grace

Once per session, you may remove a piece of armor (helm, gauntlet, or gorget) to show vulnerability. Gain +1 die to social rolls with honorable enemies, but suffer +1 Harm from the next attack against you. The mercy is real; the risk is real.

\subsubsection{The Spiral's Final Witness (10 XP)}

\textbf{Requirement:} Oath of Mercy and Grace, Tier III+

Once per session, when a creature with an active Promise Clock would deal a killing blow to an ally, you may interpose yourself. The blow is redirected to you, and you suffer the Harm instead. The Promise Clock fills immediately, and the attacker suffers the consequence (Harm, Condition, or Curse). You cannot use this talent again until you have spent a downtime scene in meditation or prayer.

\subsection{Roleplaying a Hospitaller}

\begin{itemize}
\item You do not ask which side a wounded person fought for. You bind the wound.
\item You may be called to witness deathbed confessions. The words are not yours to repeat.
\item You may struggle with anger. Mercy is not forgetfulness. The spiral remembers.
\item You may be asked to violate sanctuary—to turn someone away, to refuse the cup. You may refuse. The spiral will hold.
\end{itemize}

\subsection{Sample NPC: Sister Anabel of the Salted Veil}

\textbf{Archetype:} Hospitaller

\textbf{Patron:} Oath of Mercy and Grace

\textbf{Codex:} Grey cloak, each fold a record of a mercy given

\textbf{Thiasos:} A grey cord tied around her wrist; it tightens when she is tempted to refuse mercy

\textbf{Personality:} Calm, steady, inexhaustible. She has not slept well in years. She does not complain.

\textbf{Conflict:} A warlord she once healed now marches on the hospice where she works. She could turn him away. The spiral says no. The cord tightens.

\subsection{GM Guidance: Intrusions for Oath of Mercy and Grace}

\begin{itemize}
\item A patient the Hospitaller saved returns—not to thank, but to harm. The spiral tightens.
\item A child dies despite the Hospitaller's best efforts. The grey cord does not loosen.
\item An enemy begs for mercy. Granting it will cost the Hospitaller something real—a reputation, an ally, a future battle lost.
\item A hospice bird sings. Someone is about to die. The Hospitaller does not know who.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
