\section{Mykkiel — Arbiter of the Covenant}

\subsection{Domain and Epithets}

Mykkiel is the unwavering judge of covenants, the divine lawgiver whose scales and sword uphold sacred order. He is not good, nor evil—he is the law. Mercy without justice is weakness; justice without law is chaos. His followers are judges, templars, and scribes who enforce divine law with zealous devotion, seeing the world in absolutes of sanctity and transgression.

The Covenant, the great faith that bears his name, was revealed not through a prophet of soft words, but through a commander of hosts—a figure of burning authority who carved the first covenant into stone with a sword. Mykkiel does not negotiate. He commands. The faithful do not interpret; they obey.

\subsection{Cultural Origins}

Mykkiel appears wherever law is written in stone and enforced with steel:

\begin{itemize}
\item In \textbf{Viterra}, he is Seint Michel, the patron of knights who swear to protect the innocent and uphold the code. His shrines are courthouses and tilt-yards.
\item In \textbf{Ecktoria}, he is the Judge of the Scar, invoked before every contract. His symbol is a cracked gavel—justice that has been used.
\item In \textbf{Aeler}, he is the Stone-Judge. Dwarven arbiters inscribe his name on their gavels. His word settles disputes that could crack mountains.
\item In \textbf{Thepyrgos}, he is the Synod's patron, invoked when precedent fails and a new judgment must be carved.
\end{itemize}

\subsection{Warrior Archetype: The Justicar}

A Runic Warrior of Mykkiel is not a crusader of flame or a healer of mercy. They are a \textit{Justicar}—a lawbringer who carries the covenant in steel. Their Codex is a shield engraved with precedents or a gavel that doubles as a mace. Their Thiasos is a spirit-hound that tracks oath-breakers, a standing stone that witnesses judgments, or a sentient blade that recites the clauses of broken contracts.

\subsection{Codex Form}

The Justicar's Codex is a tool of judgment:

\begin{itemize}
\item \textbf{Shield:} A steel shield engraved with precedents. Each scar is a case won. Each dent is an appeal denied.
\item \textbf{Gavel:} A gavel carved from ironwood, its head weighted with lead. It serves as a mace in combat and a focus for judgments.
\item \textbf{Oath-Beads:} A string of carved bone or stone beads. Each bead represents a covenant witnessed. Breaking a bead voids the oath—at a cost.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Spirit-Hound of Judgment:} A hound with eyes like polished slate. It can track an oath-breaker across any terrain, and it knows when someone is lying.
\item \textbf{Standing Stone:} A small carved stone that, when placed on the ground, grows to waist height. It serves as a witness for judgments. It does not speak, but its presence is heavy.
\item \textbf{The Speaking Gavel (Sentient):} A gavel that speaks in legal citations. It can compel attention (Position +1 for one social roll) when struck against a surface.
\end{itemize}

\subsection{Core Mechanic: The Witnessed Precedent}

When a Justicar invokes a Rite that establishes a judgment or binds a covenant, they may cite a precedent from their Codex. Roll Wits + Lore (DV 3). On a success, the Rite's DV is reduced by 1 (minimum 1). On a failure, the precedent is contested—the GM gains 1 SB.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Measured Debt (4 XP)}

\textbf{Tags:} [VOW], [COMMAND], [WITNESS]

\textbf{Threshold:} A pair of scales balanced with tokens from both sides.

Establish a temporary accord. Both parties suffer -1 die if they break it first. You gain +1 die to enforce compliance.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Weighed Heart (5 XP)}

\textbf{Tags:} [TRUTH], [SENSE], [WITNESS]

\textbf{Threshold:} A small brass scale touched briefly to the chest.

Sense if the target acts against their nature or oath. Gain +1 die when pressing them. If the target is lying, the scale tips—only you can see it.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Balanced Scales (8 XP)}

\textbf{Tags:} [DEBT], [BOND], [TRANSFER], [WITNESS]

\textbf{Threshold:} A set of scales inscribed with runes of parity.

Exchange a burden between two willing parties (Harm for Fatigue, Debt for Favor, etc.). Both gain +1 die to cooperate. If the exchange is unwilling, you must win a contested Command roll (Presence + Command vs. Resolve).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Covenant Seal (9 XP)}

\textbf{Tags:} [VOW], [BOND], [DEBT], [WITNESS]

\textbf{Threshold:} A written contract and a flame to seal the wax.

Formalize a binding covenant between two or more parties. Each party must sign or mark the document. Violation of the covenant causes the violator to mark 1 Harm (Stress) immediately, and the GM gains 2 SB (the law's attention).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Final Adjudication (13 XP)}

\textbf{Tags:} [AREA], [COMMAND], [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A circle of iron bells, each etched with runes of judgment.

Convene an unseen tribunal. Shadows of former judges and wronged souls gather to preside. For the next scene, disputes within the zone are judged formally: +2 dice to Command when speaking as arbiter, and honest testimony gains +1 die. Any creature that lies within the zone suffers 1 Harm (Stress) as the tribunal's judgment falls.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Cite Precedent (4 XP)}

\textbf{Requirement:} Mykkiel, Lore 2+

Once per session, when you invoke a Rite that establishes a judgment or binds a covenant, you may declare that a past judgment covers the current situation. Roll Wits + Lore (DV 3). On a success, the Rite's DV is reduced by 1 (minimum 1). On a miss, the precedent is contested—GM gains 1 SB.

\subsubsection{Gavel's Authority (6 XP)}

\textbf{Requirement:} Mykkiel, Body 3+

Your Codex (gavel or shield) grants you authority. Once per scene, when you strike your gavel (or strike your shield with your weapon), hostile NPCs within Near range must test Resolve (DV 4) or hesitate for one exchange.

\subsubsection{Justicar's Immunity (8 XP)}

\textbf{Requirement:} Mykkiel, Tier II+

You have been granted the authority to enforce the Covenant. You cannot be arrested or detained by any lesser authority in lands that recognize Mykkiel's law. However, abuse of this immunity (committing a crime in plain sight) will cause your authority to be revoked—lose access to this talent until you undergo a ritual of restoration (a downtime project).

\subsection{Roleplaying a Justicar}

\begin{itemize}
\item You do not argue. You cite.
\item You do not forgive. You weigh. When the scales balance, the judgment is complete.
\item You may be called to witness contracts, treaties, and marriage vows. Your presence makes them binding.
\item You may struggle with mercy. The law is the law. But you have seen the law fail.
\end{itemize}

\subsection{Sample NPC: Justicar Dara of the Stone Bench}

\textbf{Archetype:} Justicar

\textbf{Patron:} Mykkiel

\textbf{Codex:} Engraved shield, each dent a case won, each scar an appeal denied

\textbf{Thiasos:} A spirit-hound that knows when someone is lying

\textbf{Personality:} Cold, precise, weary. She has never lost a case. She is not sure she has ever won one either.

\textbf{Conflict:} A covenant she witnessed twenty years ago has been broken. The breaker is her own daughter. The law says: judge. The heart says: wait.

\subsection{GM Guidance: Intrusions for Mykkiel}

\begin{itemize}
\item A covenant the Justicar witnessed years ago resurfaces. The terms were ambiguous. Both sides demand interpretation.
\item The Justicar's authority is challenged. A rival lawbringer disputes their right to judge.
\item A case the Justicar closed long ago reopens. New evidence suggests the wrong party was punished.
\item The scales tip at an inconvenient moment—not because of magic, but because the Justicar's own doubt has weight.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
