\section{Inquisitor Prime — Purity and Domination}

\subsection{Domain and Epithets}

Among zealots, witch-hunters, and templars of the Everflame, the Inquisitor Prime is venerated as the hand of absolute purity and the sword of uncompromising order. Where others see nuance, the Inquisitor sees corruption; where others see difference, they see contamination.

Those who march beneath their sigil believe that no society can stand if it tolerates the impure or the disobedient. They hunt the arcane and the aberrant—not merely to destroy them, but to bind them as slaves to order, believing even the tainted may yet serve, so long as they are shackled.

The Inquisitor Prime appears in vision as a faceless figure clad in burnished iron, with eyes like furnace-doors and a voice like grinding chains. To serve them is to embrace judgment without mercy and to live in suspicion of all who walk outside the prescribed law.

For the faithful, doubt is treason; compassion is weakness; freedom is the seed of ruin.

\subsection{Cultural Origins}

The Inquisitor Prime appears wherever order is threatened by the strange:

\begin{itemize}
\item In \textbf{Ecktoria}, the Chain-Lanterns hunt untutored magic under the Inquisitor's banner. Their lanterns burn green when corruption is near.
\item In \textbf{Thepyrgos}, the Synod's Witch-Hunter Commissariats invoke the Prime before every purge. Their writs are sealed with iron and fire.
\item In \textbf{Aeler}, the Inquisitor is called the Light of Forges. Dwarven templars kindle hearths with a brand from the temple's eternal flame—and brand heretics with the same iron.
\item In \textbf{Vhasia}, fractured cantons have each produced their own Inquisitorial orders, hunting spies, heretics, and anyone who cannot name their grandfather.
\end{itemize}

\subsection{Warrior Archetype: The Purifier}

A Runic Warrior of the Inquisitor Prime is not a judge or a healer. They are a \textit{Purifier}—a witch-hunter who carries judgment in steel and fire. Their Codex is a censer that never goes cold or a gauntlet engraved with proscription lists. Their Thiasos is a spirit-hound that smells corruption, a lantern that reveals hidden magic, or a sentient blade that has burned a hundred heretics.

\subsection{Codex Form}

The Purifier's Codex is a tool of judgment and eradication:

\begin{itemize}
\item \textbf{Censer:} A brass censer hung on three chains. It never goes cold. Its smoke reveals hidden magic (spirits, enchantments, illusions) within Near range.
\item \textbf{Gauntlet:} An iron gauntlet engraved with the names of proscribed cults and forbidden Rites. Each name is a promise kept.
\item \textbf{Brand:} A heated iron stamp. It marks the impure. It also serves as the Runekeeper's seal on writs of condemnation.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Lantern-Spirit:} A brass lantern that burns without oil. Its light reveals invisible or incorporeal creatures. It dims when the Runekeeper shows mercy.
\item \textbf{Spirit-Hound of the Hunt:} A hound with eyes like burning coals. It can track any creature that has used magic within the past day.
\item \textbf{The Censer of Unforgiving Flame (Sentient):} A brass censer that speaks in harsh, clipped sentences. Its smoke reveals lies and hidden corruption. It never forgives.
\end{itemize}

\subsection{Core Mechanic: The Hunt}

When a Purifier is hunting a specific target (a heretic, a rogue mage, a corrupted creature), they gain +1 die to all rolls related to tracking, identifying, or confronting that target. The GM names the target at the start of the hunt. If the Purifier shows mercy to the target, they lose access to this bonus for the remainder of the hunt.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Pure Flame (4 XP)}

\textbf{Tags:} [PROTECT], [RADIANT], [FEAR], [WITNESS]

\textbf{Threshold:} Ash from a burned grimoire or a sigil-scarred candle.

Gain +2 dice to resist supernatural influence (fear, charm, possession) and +1 Armor vs. magical attacks. Undead, demonic, or corrupted foes suffer -1 die against you.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Unclouded Eye (5 XP)}

\textbf{Tags:} [TRUTH], [KNOWLEDGE], [RADIANT], [WITNESS]

\textbf{Threshold:} Silver dust stirred into holy water.

Gain +2 dice to Investigation or Insight when spotting illusions, glamours, or magical deception. You cannot be surprised by sorcery this scene.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Cleansing Light (8 XP)}

\textbf{Tags:} [DISPEL], [RADIANT], [FORCE], [WITNESS]

\textbf{Threshold:} A shard of blessed mirror and consecrated oil.

Target a magical effect (ward, curse, enchantment). Test Spirit + Lore vs. DV = effect's Tier. On a hit, the effect is dispelled. On a partial, it is weakened (reduced effect or duration). On a miss, the effect lashes back—GM gains 1 SB.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Marked Prey (7 XP)}

\textbf{Tags:} [CURSE], [TRACK], [FEAR], [WITNESS]

\textbf{Threshold:} A belonging of the target mixed with blessed salt.

Mark a supernatural target: they suffer -1 die to all rolls and cannot hide with illusion. You gain +1 die to track or strike them. Undead, demonic, or corrupted foes take +1 Harm from you.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Final Admonition (13 XP)}

\textbf{Tags:} [DEATH], [BANISH], [FORCE], [FEAR], [WITNESS]

\textbf{Threshold:} Ash from a destroyed spellbook, martyr's water, and the target's true name.

Attempt to annihilate one supernatural foe. Target tests Spirit + Resolve (DV 5). On a failure, they are destroyed outright—no remains, no return. On a success, they suffer Harm 3 and -2 dice this scene. This Rite may be used only once per target.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Truth's Gaze (6 XP)}

\textbf{Requirement:} Inquisitor Prime, Presence 3+

Once per session as an action, fix your gaze on a target within Near range. For the next minute, they cannot tell a direct lie to you; they may evade or refuse to answer, but any statement they make must be factually true. Mark 1 Fatigue after using this talent.

\subsubsection{Sanctified Weapon (4 XP)}

\textbf{Requirement:} Inquisitor Prime

Once per scene, your weapon ignores 1 point of Armor from supernatural foes. The weapon glows faintly—a visible warning.

\subsubsection{Purity Zeal (6 XP)}

\textbf{Requirement:} Inquisitor Prime

When you are exposed to corruption (a curse, a demonic presence, a heretic's Rite), you may mark 1 Fatigue to gain +1 die on your next action against the source of that corruption.

\subsection{Roleplaying a Purifier}

\begin{itemize}
\item You do not negotiate with corruption. You eradicate it.
\item You may be called to witness interrogations, confessions, and executions. Your presence makes them lawful.
\item You may struggle with doubt. The Inquisitor Prime does not tolerate doubt. You hide it well.
\item You may be asked to show mercy—to a child witch, a desperate hedge-mage, a cultist who surrenders. The code says: no.
\end{itemize}

\subsection{Sample NPC: Inquisitor Voss of the Chain-Lanterns}

\textbf{Archetype:} Purifier

\textbf{Patron:} Inquisitor Prime

\textbf{Codex:} Gauntlet engraved with the names of proscribed cults

\textbf{Thiasos:} A lantern-spirit that dims when he shows mercy

\textbf{Personality:} Cold, methodical, exhausted. He has burned a hundred heretics. He remembers every face.

\textbf{Conflict:} A child has been found with a mage's knot behind her ear. The law says: burn. The child's mother is holding his hand. The lantern is dimming.

\subsection{GM Guidance: Intrusions for Inquisitor Prime}

\begin{itemize}
\item An old investigation resurfaces. Someone the Purifier once cleared is accused again—with new evidence.
\item The Purifier's zeal attracts attention. A rival Inquisitor questions their methods.
\item A supernatural entity the Purifier wounded escapes and seeks revenge.
\item The Purifier's lantern goes out at a critical moment—not because of magic, but because of doubt.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
