\section{The Witness — Truth and Revelation}

\subsection{Domain and Epithets}

The Witness remembers what others bury. Every shadow cast and oath broken is a line in her unending ledger. She is the keeper of inconvenient truths, the patron of those who seek to expose lies or recover forgotten knowledge. Her followers learn that truth comes with a price—the weight of remembering what others would forget.

The Witness has no will, only a ledger; no mercy, only accuracy. She does not demand confessions, but she ensures that every lie leaves a stain, every secret a trace. Those who serve her become archivists of the uncomfortable, witnesses to the truths that others would kill to keep hidden. To walk her path is to accept that some truths are weapons, and that the truth-teller is never loved.

\subsection{Cultural Origins}

The Witness appears wherever secrets are kept and exposed:

\begin{itemize}
\item In \textbf{Ecktoria}, she is the Ink-Eye, the Censor's clerk who copies every redacted line onto a private scroll. Her shrines are archives and scriptoria.
\item In \textbf{Thepyrgos}, she is the Fractured Mirror, prayed to by spies and scholars who seek to reveal a rival's true loyalty.
\item In \textbf{Vhasia}, she is the Fog-Scribe, whose name is whispered when a bell-line fails and the dead must be identified by their secrets.
\item In \textbf{Silkstrand}, she is the Ledger Saint, invoked before any contract is signed. Her symbol is a cracked seal—a promise that cannot be broken without witness.
\end{itemize}

\subsection{Warrior Archetype: The Truth-Sworn}

A Runic Warrior of The Witness is not a crusader or a judge. They are a \textit{Truth-Sworn}—a warrior who carries truth in steel and silence. Their Codex is a silver mirror or a locket containing a lock of hair from someone they failed to save. Their Thiasos is a raven that remembers faces, a spirit that records confessions, or a sentient blade that cannot abide a lie.

\subsection{Codex Form}

The Truth-Sworn's Codex is a record of witnessed truth:

\begin{itemize}
\item \textbf{Silver Mirror:} A hand mirror polished to a flawless sheen. It shows not the Runekeeper's face, but the face of the last person they witnessed telling a lie.
\item \textbf{Locket:} A silver locket containing a lock of hair or a written name. It is warm when the truth is near.
\item \textbf{Recording Slate:} A small slate tablet. Words written on it fade after one minute, but they can be recalled by touching the slate. Useful for capturing confessions.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Raven-Spirit:} A black bird that remembers faces. It can identify any person it has seen before, and it knows when someone is lying.
\item \textbf{Memory-Echo:} A spirit that cannot speak, but can replay sounds it has witnessed (a confession, a scream, a name spoken in the dark).
\item \textbf{The Silver Mirror (Sentient):} A hand mirror that speaks in a flat, neutral tone. It can record a single truth and replay it later. It never forgets.
\end{itemize}

\subsection{Core Mechanic: The Witnessed Truth}

When the Truth-Sworn speaks a truth that someone has tried to hide, they gain 1 Boon. This Boon can only be used on rolls involving investigation, testimony, or exposing deception.

When the Truth-Sworn is lied to, they may test Insight (Wits + Insight) vs. the liar's Deception. On a success, they know the lie. On a critical success, they glimpse the truth beneath.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Lingering Trace (4 XP)}

\textbf{Tags:} [INVESTIGATION], [KNOWLEDGE], [TRUTH]

\textbf{Threshold:} A trace of the thing to be remembered (hair, dust, a spoken name).

Gain +1 die to investigate or notice something directly related to the trace within the current scene. Ask the GM one yes/no question about whether the trace is connected to a specific person, event, or object.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Uncomfortable Question (5 XP)}

\textbf{Tags:} [COMMAND], [FEAR], [TRUTH], [WITNESS]

\textbf{Threshold:} A circle drawn with chalk or string while focusing on a specific person.

Ask one question of a target within the zone. They may either answer truthfully (the GM must give a true answer, though it may be cryptic) or mark 1 Fatigue and gain a \textit{Guilty Tell} Condition (-1 die to all social rolls until they confess the truth to someone).

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Echoing Truth (8 XP)}

\textbf{Tags:} [MIND], [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A reflective surface (mirror, still water, polished metal) used to focus on the target.

The target tests Resolve (DV 3) or suffers -1 die to rolls involving memory, deception, or resisting interrogation for the scene. If they fail, you may ask one specific, factual question about something they know, and they must answer truthfully or suffer 1 SB (Hearts) as the memory is forcibly drawn forth.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Immutable Record (7 XP)}

\textbf{Tags:} [VOW], [BOND], [TRUTH], [WITNESS]

\textbf{Threshold:} A document signed by all parties within the zone, or a spoken pact witnessed by the caster.

Bind the agreement. Any party who knowingly breaches it suffers 1 SB (Hearts) immediately and gains a persistent \textit{Oathbreaker's Mark} Condition (-1 die on social rolls involving honor, trust, or oaths until amends are made or a significant act redeems them).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Final Reckoning (13 XP)}

\textbf{Tags:} [AREA], [COMMAND], [TRUTH], [FEAR], [WITNESS]

\textbf{Threshold:} A formally called gathering (court, council, family meeting) within the consecrated zone.

All present must speak their greatest debt, wrongdoing, or hidden truth related to the gathering's purpose. Those who lie or withhold suffer Harm 2 (Stress/Reputation). Truth-tellers gain +2 dice to social actions for the remainder of the scene within the zone.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Recorded Echo (4 XP)}

\textbf{Requirement:} The Witness

Once per session, touch a physical trace (a footprint, a torn cloth, a drop of blood) and instantly know the last significant action taken by the person who left it (e.g., "they were running," "they hid something under the floor"). Mark 1 Fatigue after using this talent.

\subsubsection{Truth's Gaze (6 XP)}

\textbf{Requirement:} The Witness, Presence 3+

Once per session as an action, fix your gaze on a target within Near range. For the next minute, they cannot tell a direct lie to you; they may evade or refuse to answer, but any statement they make must be factually true. Mark 1 Fatigue after using this talent.

\subsubsection{The Scar That Testifies (8 XP)}

\textbf{Requirement:} The Witness, Witnessed Scar

Once per session, you may touch a scar and speak a single sentence about the event that caused it. That sentence is considered witnessed truth. In any court or council that recognizes The Witness, it carries the weight of sworn testimony. You cannot use this talent on the same scar twice.

\subsection{Roleplaying a Truth-Sworn}

\begin{itemize}
\item You do not lie. You may omit, you may deflect, but you do not lie.
\item You may be called to witness confessions, contracts, and deathbed testaments. Your presence makes them binding.
\item You may struggle with the weight of what you know. Some truths are weapons. Some truths are wounds.
\item You may be asked to keep a secret. The Witness does not forbid silence—only falsehood.
\end{itemize}

\subsection{Sample NPC: Archivist Venn of the Fractured Mirror}

\textbf{Archetype:} Truth-Sworn

\textbf{Patron:} The Witness

\textbf{Codex:} A silver mirror that shows the face of the last liar it witnessed

\textbf{Thiasos:} A raven-spirit that remembers every face it has seen

\textbf{Personality:} Quiet, precise, burdened. He has witnessed forty years of confessions. He remembers every one.

\textbf{Conflict:} A confession he witnessed twenty years ago was false—the confessor was protecting someone else. The true culprit still walks free. The mirror shows the liar's face every morning.

\subsection{GM Guidance: Intrusions for The Witness}

\begin{itemize}
\item The Truth-Sworn remembers something they never witnessed—a crime, a conversation, a death—in perfect detail. The Witness is sharing a memory.
\item A mirror shows not the Truth-Sworn's face, but the face of someone they wronged.
\item A secret the Truth-Sworn has kept is now known to a stranger—a blackmailer, a journalist, a rival.
\item The Truth-Sworn's Codex (mirror, locket, slate) goes blank at a critical moment. The Witness is silent. Why?
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
