\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fontenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{sectsty}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{multicol}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Aeler Module for Fate's Edge},
    pdfauthor={AI Assistant}
}

\sectionfont{\raggedright\large\bfseries}
\subsectionfont{\raggedright\normalsize\bfseries}
\subsubsectionfont{\raggedright\normalsize\itshape}

\setlist{leftmargin=*}
\setlist[itemize]{itemsep=0pt}
\setlist[enumerate]{itemsep=0pt}

\newcommand{\talent}[1]{\textbf{#1}}
\newcommand{\stringtype}[1]{\textbf{#1}}
\newcommand{\complicationtype}[1]{\textbf{#1}}

\title{Aeler: Stone, Breath, and the Ledger}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Core Concept}
The Aeler are masters of infrastructure, economics, and subtle influence. Their power lies not in flashy magic or brute force, but in their ability to make the world work - and to make others dependent on that work.

\section{Aeler — Stone, Breath, and the Ledger}

\subsection*{Mount-born engineers, keepers of underways, and masters of soft power}

\begin{quote}
``Hold what bears weight. Record what breath costs. Promise only what stone will keep.''
\end{quote}

\medskip\hrule\medskip

\subsection{I. Stone \& Breath}

The Aeler resemble their mountains: layered, load-bearing, slow to move, impossible to unmove once set. In the high Aelerian ranges, every life is counted in \emph{breaths, heat, light, and grain}. Underground, a lantern’s hue is a balance sheet; a bell-note means more than a shout. They call this discipline \textbf{deep accounting}: air tallies hung by vent-shafts, lamp-time written in chalk, calories measured in the language of ovens.

Above the tree line, stone yields little. Aeler communities rely on southern granaries to swell their numbers; in return, they build the works that make those granaries stand. To outsiders, Aeler influence feels like gravity: not a spear at the gate, but the gate itself, the levee above it, and the price of the mill-stone.

\medskip\hrule\medskip

\subsection{II. Grounds \& Holdings}

\subsubsection*{The Aelerian Range}
Alpine holds cling to ribs of granite. \emph{Vent-stacks} thrum in winter like organ pipes; ropeways hum with ore and kiln-brick. Pass-keeps mark every climb with \emph{keystone tablets}—names of those who swear to keep the road.

\subsubsection*{Underways}
A second country. Galleries where a cough wastes air; shafts where heat pools like a predator. The Aeler teach novices the \emph{Nine Measures}: light, draft, echo, dust, taste of iron, sweat-chill, lamp-shadow, bell-lag, and head-ache. Fail two, turn back. Fail three, seal the door.

\subsubsection*{The Mistlands (Protectorate)}
North of the range lies a pale country of low sun and breathing fog. Here, \textbf{Mist Wards}—slender towers with vane-bells and salt gutters—keep arable bands clear. Human farmsteads work barley, oats, and hardy roots under Aeler \emph{grain sureties}. The bargain: seed, levees, and warding for tithed surplus and winter labor on the towers. Some call it stewardship; others say the ledgers lean hard. Both are true in different winters.

\subsubsection*{Southern Commonweals}
South of the range, cities—Ecktoria, Acasia, Vhasia, Viterra, Thepyrgos, Ubral—wear Aeler work like hidden bones: bridges that never fall, floodgates that answer one key, ovens that feed thousands, mint dies that refuse to blur. Contracts are the quiet battlefield; concessions the spoils. A pledge of stone is better than a regiment that eats.

\medskip\hrule\medskip

\subsection{III. Threats from Below}

\textbf{Deep Drakes} are not simply big lizards. They are pressures that think: heat and hunger with a will, minds that push at the back of your teeth. The Aeler call the phenomenon \textbf{stone-press}. Symptoms begin as number-slips and end as people walking into shafts “to make the sums right.”

\begin{itemize}[leftmargin=*]
  \item \textbf{Signs:} condensation runs \emph{uphill}; lamp-blue goes \emph{flat}; bell-ring returns with a late second echo.
  \item \textbf{Defenses:} \emph{null-bells} tuned to kill harmonics; \emph{vein-salt} lines across thresholds; \emph{cold writs} (etched silver plates) worn over the heart; teams that count breaths in rotation and swap counters mid-phrase to catch mind-drift.
  \item \textbf{Doctrine:} never descend alone; never let a single mind do the ledgers; never meet a Drake’s gaze in still air.
\end{itemize}

\textit{In play (fronts):} \textbf{Stone-Press [6]}, \textbf{Miasma Spread [4]}, \textbf{Vent Failure [4]}. When \emph{Stone-Press} fills, escalate \emph{hallucinated ledgers} and \emph{mis-signed keystones}; treat social scenes as if the room itself is an opponent.

\medskip\hrule\medskip

\subsection{IV. Law, Ovens, and Soft Power}

Aeler law is \textbf{Tally-Law}: if it isn’t written, it isn’t owed; if it cannot bear weight, it isn’t promised. Their soft power rides four pillars:

\begin{itemize}[leftmargin=*]
  \item \textbf{Keystone Rights.} Control the piece that holds the whole. Aeler sign \emph{keystone clauses}: maintenance for access, repair for rate. Break the clause, and the bridge “politely rests.”
  \item \textbf{Grain Surety.} Winter ovens and storage domes. In exchange for tithe windows and public pricing courts, cities adopt \emph{Oven Charter Days}—no one starves while the charter bell rings.
  \item \textbf{Water \& Flood.} Sluice math is power. Aeler sluicewarden guilds lease keys, not walls; they can open a city for trade faster than any army can take it.
  \item \textbf{Mint \& Measure.} Calibrated weights, indelible dies. When a coin bears Aeler marks, courts sharpen.
\end{itemize}

\textit{In play (Strings):} \emph{keystone right}, \emph{grain surety}, \emph{sluice key}, \emph{mint die}, \emph{oven charter}. Cash a String to shift a negotiation’s DV by --1 or to force a \emph{public bowl} (fairness hearing) before a blade is drawn.

\medskip\hrule\medskip

\subsection{V. People \& Institutions}

\begin{itemize}[leftmargin=*]
  \item \textbf{Bell-Chains.} Signal corps that read echo and carry news down-lines faster than runners. Their captaincy exams are held in the dark.
  \item \textbf{Ledger-Kin.} Families who hold communal contracts and assign \emph{scrips of air, heat, and lamp}. Marriages are audited as carefully as bridges.
  \item \textbf{Vent Priors.} Engineers who treat airflow like liturgy. Their vestments are leather aprons burned in a pattern that maps the hold.
  \item \textbf{Oven-Wardens.} Bakers as quartermasters; saints of bread and ration-line. In famine, they outrank colonels.
\end{itemize}

\medskip\hrule\medskip

\subsection{VI. Aesthetics \& Speech}

\begin{itemize}[leftmargin=*]
  \item \textbf{Colors:} soot, iron, kiln-orange, glacial blue.
  \item \textbf{Materials:} rough-split stone outside; mirror-polished keystones within; leather scored with counting notches; vellum that smells of lye and smoke.
  \item \textbf{Proverbs:} 
  \begin{itemize}
    \item Stone keeps the promises you place on the right spot.
    \item Count in the light, breathe in the turn.
    \item A bridge is a treaty that learned to carry weight.
    \item Heat is a loan; pay it back with bread.
  \end{itemize}
  \item \textbf{Names:} given + craft + keep: \emph{Orra Vent-Prior of Third Stack}; \emph{Ghent Bell-Captain of Red Echo}. Silence is punctuation; bells are syntax.
\end{itemize}

\medskip\hrule\medskip

\subsection{VII. Relations \& Rivalries}

\begin{itemize}[leftmargin=*]
  \item \textbf{Ykrul.} Mutual respect where work meets weather. Rope meets ring; pass-oaths meet meadow concessions. Arguments end on the Board with a hand on the stone.
  \item \textbf{Linns.} Sea people who love a keystone you can sail; Aeler admire their hull-geometry, distrust their ledgers.
  \item \textbf{Lethai-al.} Tree-law and stone-law debate. Root and ridge agree more often than songs admit.
  \item \textbf{Deep Drakes.} The foe that teaches humility. Victories are written in sealed doors and names of the lost.
\end{itemize}

\medskip\hrule\medskip

\subsection{VIII. Strings \& Tools (at the table)}

\begin{itemize}[leftmargin=*]
  \item \textbf{Keystone Tablet.} Establishes or pauses a route; once/scene, convert a chase into a stand where \emph{Position +1} for defenders who prepared.
  \item \textbf{Null-Bell.} On ring, cancel one \emph{psychic push} or compel; costs \emph{lamp-time} (mark gear wear or Fatigue).
  \item \textbf{Oven Charter Seal.} Force a \emph{public bowl} in markets: one round of fair-price negotiation before violence can escalate.
  \item \textbf{Sluice Key.} DV --1 on any operation hinging on water, flood, or sanitation; if abused, create \emph{Public Outrage [4]}.
  \item \textbf{Air Scrip.} Negate a first underground suffocation/miasma consequence in a leg; on use, tick \emph{Vent Failure [1]}.
\end{itemize}

\medskip\hrule\medskip

\subsection{IX. Adventure Seeds}

\begin{enumerate}[leftmargin=*]
  \item \textbf{The Bell That Rang Twice.} A late echo signals stone-press in the east galleries. Audit the vent ledgers, find the mind-heat, and seal what thinks in the dark.
  \item \textbf{Bread Before Swords.} Southern grain is short; an Oven Charter is contested by syndics with sharp friends. Win the bowl, post the rate, keep the line fed.
  \item \textbf{Keystone Courtesy.} A bridge fails “politely” on the caravan road. Someone altered the tablet. Follow the chisels to a quiet war over a toll.
  \item \textbf{Mist Ward Winter.} Warding bells are silent and the fog eats fences. Repair the tower with farm help who distrust the ledger—and prove the sums keep them alive.
  \item \textbf{Drake’s Ledger.} A Deep Drake hoards \emph{numbers}: stolen tallies, miscounted breaths, lost names. Bring back the books and the people they keep.
\end{enumerate}

\medskip\hrule\medskip

\subsection{X. Portraying the Aeler}

Lead with craft and capacity. Show how a promise is a structure, not a sentence. Underground, make \emph{air and light} into currencies; overhead, let contracts move armies without banners. When the deep presses, slow your words, count your breaths, and let the bell decide who speaks next.

\medskip\hrule\medskip

\subsection{XI. Sumwrights \& Ledger-Compacts (Aelinnel Allies)}

Aeler holds commonly retain \textbf{Aelinnel Sumwrights} as neutral keepers of ledgers. Copper-stylus and split-book in hand, they maintain the \emph{two-ledger doctrine} (said/meant) and audit ovens, sluices, and mints. Sumwright contracts insist on:
\begin{itemize}[leftmargin=*]
  \item \textbf{Public Weights.} Scales, measures, and dies are tested at market-bell; falsifiers lose \emph{Display Rights} (see XII) for a season.
  \item \textbf{Even Witness.} A Sumwright must read accounts to both parties aloud; silence equals assent for that line only.
  \item \textbf{Split Custody.} One ledger remains in the hold; one travels with the Sumwright caravan; discrepancies ring the \emph{Red Bell}.
\end{itemize}
\textit{Strings:} \emph{sumwright retainer}, \emph{public weight right}, \emph{red-bell audit}. Cash to reduce a negotiation DV by --1 or force a fairness hearing before a blade is drawn.

\medskip\hrule\medskip

\subsection{XII. Commerce, Status, and Display}

Aeler society is \textbf{collectivist and mercantile}: wealth flows first to commonworks (ovens, wards, roads, keystones), then returns as \textbf{Display Rights} that signal status. The more you contribute, the more you may wear, carve, or light.
\begin{itemize}[leftmargin=*]
  \item \textbf{Display Charter.} Public marks (metal trim, lamp-halos, keystone etching) are licensed by the ledger-kin. Fraudulent display is a civic shame and fined in grain or labor.
  \item \textbf{Rings of Account.} Narrow bands on tools and belts show contract classes kept: water, bread, bridge, mint. Three rings grant first voice in oven courts.
  \item \textbf{Quiet Wealth.} Private hoard without public work is suspect; \emph{unworked shine} draws auditors and jokes in equal measure.
\end{itemize}
\textit{At the table:} presenting a valid \emph{Display Writ} improves \emph{Position +1} in civic negotiations once/scene; on a miss, start \emph{Audit Clock [4]}.

\medskip\hrule\medskip

\subsection{XIII. Orders \& Companies}

\begin{itemize}[leftmargin=*]
  \item \textbf{Iron Avengers.} Traditionalists who enforce blood-feuds inscribed on keystones. They keep \emph{Gray Lists} (names owing red silver) and are bound to answer when rung. Most cities tolerate them only under license.
  \item \textbf{Spirit Shield Warriors.} Ancestor-venerating protectors who wear \emph{mask-helms} etched with lineage prayers. They guard vent-stacks, memorial galleries, and children’s routes. When the \emph{Stone-Press} rises, they stand first.
  \item \textbf{True Masons.} Wanderers of Acasia, Vhasia, and Viterra who repair ancient Aeler work—bridges, sluices, ovens, wards. Paid in tithe-windows, charter clauses, and a bed in any house that stands because of their hands.
  \item \textbf{Edgewalkers.} Opportunists who scan borders and markets for the gap that pays. Part scout, part broker, part thief-catcher, they sell \emph{first knowledge} of breaks and bargains.
  \item \textbf{Reform Lodges.} Aeler who object to human exploitation concentrate in foothill freeholds and human cities. They press for \emph{grain sureties without hooks}, teach oven math to neighbors, and vote their ledgers in favor of mixed councils.
\end{itemize}

\textit{Strings \& Tools:} \emph{gray list token} (compels a parley before feud blows), \emph{ancestor mask} (once/scene resist fear/psychic pressure), \emph{mason’s oath} (DV --1 on repair/restoration scores), \emph{edgewalker marker} (create an opportunity clock on first contact).

\medskip\hrule\medskip

\subsection{XIV. Ethics \& Dissent}

Many Aeler reject exploitative arrangements with human protectorates. You find them teaching weights in river towns, serving as oven-wardens in mixed boroughs, and organizing \emph{Charter Kitchens} where the bell rings regardless of ledger. In Mistland winters, these dissenters argue for \emph{tithe holidays} and push to convert ward-labor into paid guild work. Their opponents answer with ledgers; their allies answer with bread.

\medskip\hrule\medskip

\subsection{XV. Play Hooks (Orders in Motion)}

\begin{enumerate}[leftmargin=*]
  \item \textbf{Red Bell, Gray List.} An Iron Avenger ring forces a feud claim in a city that banned duels. Can you translate blood into bread-price and retire the name from the stone?
  \item \textbf{Masks in the Miasma.} Spirit Shields request outside help to clear a vent gallery where Stone-Press twists numbers. Keep the count, keep your mind.
  \item \textbf{Bridge Without Purse.} True Masons will repair a flood-broken span if a mixed council signs a Display Charter that elevates commonworks over private shine.
  \item \textbf{Edge of Profit.} An Edgewalker offers first knowledge of a Syndicate price shock at the harbor. Take the lead, share the gain, or watch the sluice open for someone else.
  \item \textbf{Charter Kitchen.} Reform Lodges move to open a winter oven without tithe. Broker grain, tame politics, and post the bell before the ledgers close the door.
\end{enumerate}

\section*{Affinity (Aeler)}
\begin{itemize}
  \item \textbf{Stone \& Breath:} +1 die to Craft, Tinker, and Survival rolls related to infrastructure, construction, or resource management. In underground/dense urban environments, +1 Position on rolls to navigate, maintain, or sabotage systems.
  \item \textbf{Deep Accounting:} Once per scene, you can spend 1 Boon to ``audit'' a situation - ask one question about hidden resources, costs, or dependencies that the GM must answer truthfully.
\end{itemize}

\section*{Starting Talent}
\talent{Vent Prior's Training (3 XP - Minor Talent):}
\begin{itemize}
  \item Requirements: Craft 1+, Wits 2+
  \item Benefits: 
    \begin{itemize}
      \item +1 die to rolls involving air quality, ventilation, structural integrity, or underground navigation.
      \item Know the ``Nine Measures'' (light, draft, echo, etc.) - can detect environmental hazards or hidden passages with a successful Wits + Notice roll (DV 3).
      \item Once per scene, can ``read'' a structure like a ledger, gaining +1 die to rolls to understand its construction, weaknesses, or maintenance needs.
    \end{itemize}
\end{itemize}

\section*{Key Talents}

\subsection*{Mint \& Measure Path}
\talent{Calibrated Weight (4 XP - Minor Talent):}
\begin{itemize}
  \item +1 die to rolls involving trade, negotiation, or detecting counterfeits.
  \item Once per scene, can force a ``public bowl'' (fairness hearing) in a market or negotiation scene by spending 1 Boon. This shifts the DV of the negotiation by -1 and gives you +1 Position.
\end{itemize}

\talent{Keystone Clause (6 XP - Major Talent):}
\begin{itemize}
  \item Requirements: Calibrated Weight
  \item Benefits: 
    \begin{itemize}
      \item Can establish ``keystone rights'' over a piece of infrastructure, contract, or resource. This gives you a String (see below).
      \item Once per scene, can ``pause'' a route, contract, or service you control, converting a chase into a stand (Position +1 for defenders who prepared) or forcing a negotiation.
    \end{itemize}
\end{itemize}

\subsection*{Oven-Warden's Path}
\talent{Charter Bell (5 XP - Minor Talent):}
\begin{itemize}
  \item Requirements: Presence 2+, Command 1+
  \item Benefits:
    \begin{itemize}
      \item In a crisis involving survival (famine, shelter, etc.), can organize communities with a successful Presence + Command roll (DV 3). Success grants +1 Position to all allies for survival-related actions.
      \item Once per scene, can invoke an ``Oven Charter'' - force a public bowl in a crisis situation, ensuring fair distribution of resources or preventing violence.
    \end{itemize}
\end{itemize}

\subsection*{Vent-Captain's Path}
\talent{Bell-Chain Signal (4 XP - Minor Talent):}
\begin{itemize}
  \item Requirements: Wits 2+, Notice 1+
  \item Benefits:
    \begin{itemize}
      \item +1 die to rolls involving communication, signals, or detecting hidden information.
      \item Can establish a ``bell-chain'' communication network with allies. Once per scene, can send a message instantly to any ally within the network, regardless of distance (as long as there's a path).
    \end{itemize}
\end{itemize}

\section*{Strings (Aeler)}
Aeler characters accumulate ``Strings'' - markers of influence and control over infrastructure, contracts, and resources. These work like Bonds but are more concrete and economic.

\begin{itemize}
  \item \stringtype{Keystone Right:} Control a critical piece of infrastructure (bridge, sluice, ventilation shaft). Can ``pause'' it once/scene.
  \item \stringtype{Grain Surety:} Control a food source or storage. Can force fair distribution or gain leverage in negotiations.
  \item \stringtype{Sluice Key:} Control water flow or sanitation. Can shift DV by -1 on operations involving water/flood.
  \item \stringtype{Mint Die:} Control a source of currency or measurement. Can force a public bowl or detect counterfeits.
  \item \stringtype{Oven Charter:} Control a public resource (oven, shelter, etc.). Can invoke charter protections.
\end{itemize}

\section*{Patron Relationships}
Aeler typically bond with Patrons related to \textbf{Order, Infrastructure, Wealth, or Craft} (Sacred Geometry, Clockwork Monad, Mab, The Witness, Oath of Flame \& Light). Their approach to patronage is often contractual - they make deals and expect precise repayment.

\section*{Complications}
\begin{itemize}
  \item \complicationtype{Stone-Press Susceptibility:} When in deep underground environments or high-stress situations, make a Spirit test (DV 3) or suffer -1 die to rolls due to hallucinations or mental pressure.
  \item \complicationtype{Ledger Dependency:} You rely on precise accounting. If separated from your ledgers/records for more than a day, suffer -1 Position on rolls involving planning or resource management.
  \item \complicationtype{Surety Obligation:} You've guaranteed resources to a community. If you fail to deliver, mark 2 segments on your Obligation clock with your primary Patron (or gain a permanent Complication if you don't have one).
\end{itemize}

\section*{Sample Aeler Character}

\textbf{Orra Vent-Prior of Third Stack}
\begin{itemize}
  \item Body 2, Wits 3, Spirit 2, Presence 2
  \item Skills: Craft 2, Notice 2, Survival 1, Tinker 1
  \item Talents: Vent Prior's Training, Bell-Chain Signal
  \item Affinity: Stone \& Breath
  \item Strings: Keystone Right (on the main elevator shaft to the surface), Sluice Key (controls the water reclamation system)
  \item Complication: Stone-Press Susceptibility
\end{itemize}

\section*{Adventure Hooks}
\begin{enumerate}
  \item \textbf{The Bell That Rang Twice:} Investigate a malfunctioning signal system in an Aeler hold that's causing dangerous delays. Is it sabotage, a Deep Drake, or something worse?
  \item \textbf{Bread Before Swords:} A city's Oven Charter is being contested by a powerful syndicate. Can you ensure the winter ovens stay open and fair?
  \item \textbf{Keystone Courtesy:} A bridge has ``politely rested'' - someone altered the keystone tablet. Track down the culprit through the trail of chisel marks and contract disputes.
  \item \textbf{Mist Ward Winter:} The Mistland wards are failing and the fog is eating the farms. Repair the towers with locals who distrust Aeler ``ledgers'' - and prove the math keeps them alive.
  \item \textbf{Charter Kitchen:} A Reform Lodge wants to open a winter oven without tithe. Broker grain, tame politics, and post the bell before the ledgers close the door.
\end{enumerate}

\section*{Integration with Core Rules}

\subsection*{Deep Drakes as Fronts}
When encountering Deep Drakes, use the provided fronts:
\begin{itemize}
  \item \textbf{Stone-Press [6]}
  \item \textbf{Miasma Spread [4]}
  \item \textbf{Vent Failure [4]}
\end{itemize}
When Stone-Press fills, escalate with hallucinated ledgers and mis-signed keystones.

\subsection*{Tally-Law Mechanics}
Aeler law (Tally-Law) can be represented through:
\begin{itemize}
  \item Using Strings system for contractual obligations
  \item Applying -1 DV in negotiations when proper documentation exists
  \item Creating clocks for ``Audit [4]'' when contracts are violated
\end{itemize}

\subsection*{Display Rights}
Aeler status through Display Rights can be tracked as:
\begin{itemize}
  \item Minor Asset (4 XP): Public Display Charter - +1 Position in civic negotiations
  \item Major Asset (8 XP): Rings of Account - First voice in oven courts, DV -1 on resource management
\end{itemize}

\end{document}


