\documentclass[12pt]{article}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,      
    urlcolor=cyan,
}
\title{The Book of Mist, Air, Alder, Thorn, and Mirror}
\author{Scribe Valerius of the Thepyrgos Synod\\
Special Commission for the Study of Indigenous Customs}
\date{Presented to the Honorable Synod in the Year 847 AR,\\
under the seal of the Archon's Third Clerk.}

\newpage

\begin{document}
\maketitle

\input{copyright}
\section{Four Investigations into the Threshold Cultures of the Northern and Eastern Shores}
\textit{Compiled by Scribe Valerius of the Thepyrgos Synod, Special Commission for the Study of Indigenous Customs, with annotations, marginalia, and a growing conviction that the author has spent too long in the field.}
\textbf{Presented to the Honorable Synod in the Year 847 AR, under the seal of the Archon's Third Clerk.}

\subsection{A Note on Method and Madness}
\textit{It has come to my attention that the Synod expects a straightforward ethnographic report. I regret to inform the Honorable Body that the subject matter resists straightforwardness as stone resists water -- that is to say, eventually, but only after wearing the investigator down to smoothness.}

The following pages contain four tales, each drawn from a distinct threshold culture of the northern and eastern shores. They are not folk tales in the Thepyrgosian sense -- those tidy narratives with clear beginnings, middles, and instructive endings. They are, rather, \textit{accounts of debts}: debts owed to stone, to fog, to root, and to the silence between counted breaths.

I have recorded them as faithfully as my training permits. I have added footnotes where the original defied translation. I have suppressed the urge to add a conclusion, because every attempt at a conclusion was contradicted by the next informant, and the next, and the next.

\textit{Valerius of Thepyrgos}\\
\textit{Payden's Port, in the third week of fog}\\
\textit{The lamp is guttering. The ink is cold.}

\section{Part One: The Keystone That Counted Three}
\subsection{An Aeler Tale from the Vault‑Kingdom of Khaz‑Vurim}
\subsubsection{As told by Vault‑Brewer Hetha, who lost her voice to a gas‑flare and speaks now by tapping her chisel against her teeth. Recorded in the under‑crypt of the Third Deep.}

The Aeler of Khaz‑Vurim do not lose things. They \textit{misplace them by a factor that can be corrected with interest}.

This is an important distinction. A loss is a vacancy in the world. A misplacement is merely a temporary asymmetry -- an error in the ledger that can be balanced, given enough time and the correct application of arithmetic.

When the Vault‑Queen's favoured engineer, one \textbf{Dorin Keystone‑Scribe}, disappeared into the workings of the Seventh Shaft, the ledgers showed no anomaly. He had been allotted eight hundred breaths of air, a lamp‑time of five bells, and a lunch of salt‑biscuit and pickled fungus. All were consumed. The ledger closed.

\footnote{The Aeler concept of "closure" differs from our own. A Thepyrgosian notary closes a file when all parties are satisfied. An Aeler closes a file when the numbers balance, regardless of whether any party is present to register satisfaction. This is, I am told, not cruelty. It is efficiency.}

But the vent‑priests knew better. The airflow in the Seventh Shaft had shifted -- not much, a thumb's width on the pressure‑gauge, but the Deep Watchers count in \textit{hairs}. Something was breathing down there that should not have been breathing.

They sent Correctors. The Correctors did not return. \footnote{Their names have been redacted from the public ledger. Their breath‑allotments, however, remain on the books, accruing interest.}

They sent a Spirit‑Shield. The shield returned empty, its copper rings corroded, its bell‑chime a flat note that does not exist in any scale known to the Synod's acousticians. I have heard a recording. It is reminiscent of a door closing under water.

Finally, they sent an Edgewalker -- a soft‑exiled scout named \textbf{Rinn of the Broken Trace}, who had been reassigned to the surface for thirty years. The Aeler do not exile. They \textit{reassign}. The Mistlands, the Edgewalker companies, the far salt‑pans -- these are the destinations of those whose presence in the deep holds has become, for one reason or another, \textit{inconvenient}.

\footnote{Rinn's original offense was questioning a keystone clause in the Third Vent's air‑allotment ledgers. The Prior was offended. The Prior had cousins. This pattern is, I suspect, universal.}

Rinn took a slate of bleached quartz, a null‑bell calibrated to the negative harmonics of the seventh depth, and a single Obsidian chipped from the Vault‑Queen's own seal‑ring. She descended the Seventh Shaft.

The shaft had changed. It was not deeper -- Khaz‑Vurim's depths are measured in generations, not meters -- but the \textit{angles} were wrong. A corridor that should have led east now bent west. A stair that should have had twelve risers had thirteen. The thirteenth riser was warm.

Rinn counted her breaths. One, two, three, four, five, six, seven, eight. Then she stopped. The Aeler do not count nine in the deep. \footnote{This prohibition recurs across multiple cultures of the northern and eastern shores. The number nine appears to be a boundary marker between the countable and the not‑countable. I have collected seventeen variants of this taboo. None of them conclude satisfactorily.}

A voice spoke from the warm step. It was Dorin's voice, but Dorin had been dead for a year. I am aware that the Skeptical Reader will question the veracity of a dead man speaking. I can only report what I was told.

\textit{``I mis‑counted the keystone," said the voice. "By three degrees. The vault is holding, but the interest is due."}

Rinn did not flee. Edgewalkers are not trained to flee; they are trained to \textit{measure} the distance to the nearest exit and then decide, with grim precision, whether fleeing is worth the breath.

\textit{``What is the interest?" she asked.}

\textit{``Three names. One for each degree. I gave one. My wife. She coughs now, in the dry season. I will give the second. My child. He will dream of falling. The third -- the third is the Vault‑Queen"s. But I cannot take her name. She is not mine to owe."}

\footnote{The Aeler concept of "owing" is distinct from our own. A debt is not a moral failing. It is a \textit{structural feature}. The mountain owes its shape to the force that lifted it. The keystone owes its position to the pressure of the arch. To owe is to be in proper relation. To owe incorrectly is to be a cave‑in waiting to happen.}

Rinn understood. Dorin was not a ghost. He was a \textit{ledger error} -- a debit that had not been offset, now accruing compound interest in stone.

\textit{``What do you need, Keystone‑Scribe?"}

\textit{``Find the keystone. The one I mis‑counted. Tap it three times -- once for each degree -- and speak the True Measure. The rock will remember. Then I can stop counting."}

Rinn descended deeper. The null‑bell rang once -- a warning that something was pressing against the geometry from the side of the ledger that has no columns.

She found the keystone in a chamber that had not existed when Khaz‑Vurim was first carved. The chamber was featureless except for the stone itself: a block of black granite, warm to the touch, its face inscribed with nine tally‑marks. The ninth mark was still being carved -- not by a chisel, but by the slow accretion of dust, one grain at a time, as if the stone were counting down from an enormous number and had not yet reached the bottom.

Rinn tapped the keystone three times -- one, two, three -- each tap spaced by a single breath.

\textit{``True Measure," she said, "is not eight. True Measure is \textit{enough}."}

\footnote{I have spent three months trying to verify whether "enough" is a technical term in Aeler metrology. The consensus among my informants is that it is not, but that it \textit{should be}, and that Rinn's use of it was either inspired or heretical. The Vault‑Queen declined to clarify.}

The ninth tally‑mark stopped growing. The stone exhaled -- a sound like a mountain settling after a long argument, or a ledger being closed with a heavy hand. The corridors realigned. The thirteenth riser became the twelfth. The null‑bell fell silent.

Rinn returned to the surface. Dorin's name was moved from "missing" to "closed". His wife's cough ceased. His child's dreams were, by all accounts, his own again.

The Vault‑Queen summoned Rinn. \textit{``What did you learn?" she asked.}

Rinn said: \textit{``We count to eight because nine is a door. But sometimes, the door is not an enemy. Sometimes it is a corridor that has been waiting for the right measure."}

The Vault‑Queen did not smile. Aeler queens do not smile in public -- smiling is, I am told, a form of giving, and the Aeler are careful with gifts. But she did not reassign Rinn back to the surface. The Edgewalker kept her deep.

\footnote{I asked Vault‑Brewer Hetha what became of Rinn. She tapped her chisel against her teeth in a pattern I have since learned means "that is a different story, and you have not paid for it." I paid. The story was long. I will summarize: Rinn eventually became Warden of the Seventh Shaft. She still counts in eights. The ninth door, she says, is still there -- but now it opens for her, and only for her, and only on days that do not have names.}

\subsubsection{The Scholar's Observation}
\textit{It is tempting to read this tale as a parable about the danger of precision without wisdom. I suspect the Aeler would find such a reading reductive. The danger, from their perspective, is not precision -- precision is the only thing that keeps the mountain from falling -- but \textit{unchecked} precision, arithmetic without mercy, count without enough.}

\textit{The keystone remembered the true number because it was designed to. The error was human. The correction was also human -- but only because Rinn was willing to speak a word that did not appear in any ledger.}

\textit{``Enough" is not a number. It is a decision. The Aeler do not have a word for "enough" that is not also a word for "the sound a bell makes when it is rung by a hand that intends to stop." I have heard that sound. It is quieter than you would think.}

\section{Part Two: The Bell That Does Not Tire}
\subsection{A Mistlands Tradition, as told by Warden‑Inspector Brom, recorded at Payden's Port}
\emph{Editor's Note: This account was given in fragments over three nights of salt tea and silence. The following is a synthesis. Any errors are mine. The fog, as they say, has a way of editing.}

The Mistlands are not a place. They are a \textit{process}.

This is the first thing Warden‑Inspector Brom said to me. It was also the last thing he said, each night, before his silence grew too heavy for questions.

The process, as I understand it, is as follows: bells keep the Direwood from walking. Salt keeps the dead from remembering their names. The fog keeps everything else at a distance -- including certainty.

Brom was an Edgewalker, originally -- he had been reassigned from the Aeler holds for questioning a keystone clause, or perhaps for failing to question one; the details varied with his mood and the strength of the tea. He had served in the Mistlands for forty‑three years. He had never rung the ninth bell.

\footnote{The bell‑line at Payden's Port consists of eight bells to a span. The ninth bracket is empty. The rope is still there. It is never touched, but it is never cut. I asked why. Brom said: "Because cutting it would be a decision. The Mistlands prefer uncertainty. It gives them something to lean on."}

When I asked him about the Warden Who Forgot to Count -- a figure that appears in several Mistlands traditions -- he fell silent for an hour. Then he spoke.

I was twenty‑two. Fresh from the holds. Still counting in my head -- breaths, steps, the number of rivets on my belt‑buckle. The Prior had sent me to Payden's Port because I had asked a question that did not have an answer, and she was tired of my asking.

The old warden met me at the quay. His name was Torvin. He had a beard that would have shamed a mountain and a left hand missing three fingers. The remaining two held a clay cup of something that smelled like fish and death.

\textit{``What'd you do?" he asked.}

I told him.

He laughed. It was the sound of a rockfall that had been waiting for an excuse.

The first week was silent. Torvin did not speak unless the fog changed. I learned to read his silences -- the way the fog turned from gray to lead, the way the bells answered when the tide turned, the way the salt pans crackled when something was pressing against the ward from the other side.

\footnote{I asked Brom what "the other side" meant. He said: "The other side of the count. Every number has two sides. Eight is the number of the living. Nine is the number of the door. The thing that presses against the bell is the number between -- the one we forgot to name." I have since learned that the Lethai‑al have a word for this gap. It is also their word for "the moment before a lie becomes true."}

On the fifth night, we came to the broken pier. Payden's Port is not a harbor. It is a collection of buildings that have given up on being dry, arranged in a crescent around a pier that leads nowhere -- the seaward end collapsed forty years ago, and no one had rebuilt it because the sea remembers whose fault it was.

The tide turned at midnight. Torvin refused to come. \textit{``Some debts," he said, "are not mine to witness."}

I stood on the broken pier alone. The fog was thick as wool. The water lapped at the pilings with the sound of a tongue trying to remember a word.

Then I heard it.

A bell. Not iron. Not bronze. Something older -- stone, maybe, or bone. It rang once. The note was not any key I knew. It was the note of a door closing when no door was there.

It rang again. Twice. A pause. Then again. Thrice.

I counted. One, two, three, four, five, six, seven, eight.

The ninth did not ring.

But I heard it anyway.

\footnote{Brom would not describe the sound of the ninth bell. He said: "Describe color to a person born in the fog. You can"t. But you know it when you see it -- or hear it, or fail to hear it, which is the same thing in the Mistlands."}

I turned. The pier was longer than it had been. At the far end stood a figure in a warden's cloak -- but his face was not a face. It was a bell. A bell with a clapper made of bone, and the clapper was moving, and the bell was not ringing because the clapper was striking \textit{silence}.

The bell‑face raised a hand. The hand had no fingers -- only iron nails driven into the palm, each stamped with a name I could not read.

It pointed at me.

Then the fog closed.

I was alone on the short pier. The water lapped. The ninth bell did not ring again. But I have been counting ever since.

\subsubsection{The Scholar's Observation}
\textit{I asked Brom if the Warden Who Forgot to Count is a ghost. He said the word "ghost" is too small -- it implies a single death, a single debt, a single mis‑count that can be corrected and closed. The Warden is not a ghost. The Warden is a \textit{reminder}, carved into the fog, that the ninth bell is always a possibility and never an event.}

\textit{I asked if the Warden appears to everyone. Brom said: "Everyone who is new to the Mistlands. Everyone who has not yet learned that counting is not the same as hearing. The fog takes time to know you. The Warden takes less."}

\textit{I have been in the Mistlands for six weeks. I have not seen the Warden. I am not sure if this means that the fog does not know me, or that it knows me too well to bother.}

\section{Part Three: The Root That Knotted Itself}
\subsection{A Lethai‑al Tale of the Hazel Court and the Alder King}
\subsubsection{As told by a pathweaver who refused her name, recorded in the twilight tongue and rendered into Utaran script with lamentation.}
\textit{The Scholar's Prefatory Lament: The Lethai‑al have no word for "moral." They have a word that signifies "the shape the axe leaves in the wood after the axe is withdrawn, and also the sound the tree does not make, and also the color of bread when you have forgotten hunger." I have rendered this concept as "lesson," though I do so with the full knowledge that my successors will find it wanting.}

In the days before the Star‑Roads were broken -- when the roots of the Valewood yet remembered the names the First Trees spoke to one another across the spaces between the world's waking -- there dwelt a pathweaver of the Hazel Court whose name in the high tongue was \textit{Serevindë}.

\footnote{The name resists translation. It contains the concept of "a raven's decision to turn its head," which is to say, a choice that has not yet announced itself. I leave it untranscribed rather than shame myself with approximation.}

Her sister, with whom she shared the grove, was called \textit{Lúvandil} -- "Daughter‑of‑the‑Breath‑that‑Smells‑of‑Rain." This translation is adequate. The rain in question is particular to the western slopes of the Valewood during the third week of autumn, when the copper‑beech leaves have turned but not fallen. I mention this because it becomes important later.

The sisters held the grove together until Serevindë coveted the western slope. She did not ask for it. She did not take it. She \textit{omitted} it -- from every accounting, every harvest‑song, every naming of trees at the turning of the season.

\footnote{The Lethai‑al verb "to omit" is also the verb "to breathe" in the imperative mood. One omits as one breathes -- without intention, without cessation, without apology. The Synod's linguistics committee has requested a full paradigm. I have declined.}

Lúvandil said nothing. She took her grievance to the roots, not to the court, and the roots remember what the court forgets. Then she died -- not of the omission, but of a fever brought by the lowland Men.

The grove passed entire to Serevindë. But the blight came. Bark‑fungus with faces. The faces were not angry. They were \textit{hungry} -- the hunger of things that have been given no name.

Serevindë walked the star‑roads to the Hazel Queen. \textit{``Lift the blight," she said, speaking in the Old Tense that binds future to past and past to the interest that accrues in between.}

The Hazel Queen said: \textit{``Tell me a story I have never heard."}

Serevindë told a story -- true as far as it went. She did not speak of Lúvandil. A story without a sister is still a story, she told herself. The Hazel Queen knows that a story without a sister is a debt accruing interest.

\textit{``The blight will lift," said the Queen. "But you omitted the Daughter‑of‑Rain‑Breath. Her ghost now walks the grove at dusk. I did not agree to lift the ghost."}

So it was. The fungus faded. But at dusk, Lúvandil walked between the oaks -- doing nothing, demanding nothing, simply \textit{being} where she had been omitted. A ghost that does nothing is worse than a ghost that rages, because nothing cannot be answered.

\footnote{The Lethai‑al have a specific term for this kind of haunting: \textit{mornínë}, which translates approximately as "the presence of a debt that has not yet decided what it wants."}

Serevindë returned to the Hazel Court. \textit{``The ghost remains," she said.}

\textit{``You did not ask for its removal," said the Queen. "Now you owe me two stories, and you must tell them in the presence of the Alder King."}

The Alder King's hall is a hollow log that has no entrance -- except that you are already inside when you realize you should have knocked. His voice is frost splitting granite.

\textit{``Tell," he said.}

Serevindë told the true story, with Lúvandil's name woven into every clause, every subclause, every subordinate conjunction that the Lethai‑al possess (and they possess many, for their grammar is the grammar of things that do not wish to be misunderstood). She did not ask for forgiveness. The Alder King does not grant forgiveness. He grants \textit{reckoning}.

\textit{``The ghost remains," he said, "because you still owe a gift. She gave you the land freely. You gave her nothing. Omission is a hunger. The ghost is hungry. It will not eat until you give it something it can swallow."}

\textit{``What do ghosts swallow?" asked Serevindë.}

\textit{``Names," said the Alder King.}

But Serevindë had forgotten Lúvandil's name. The omission had eaten it -- slowly, the way water eats stone, the way a story without a sister eats itself from the inside.

She returned to her grove. At dusk, she sat beneath the copper‑beech and began to list everything she remembered -- not the name, but the shape of her sister's hands, the sound of her laugh when the cider soured, the scar on her left knee from a fall on the fallen Star‑Road. She spoke for three dusks.

On the fourth, the ghost opened its mouth. A single breath, smelling of apples and iron.

\textit{Lúvandil.}

The name.

Serevindë wept -- a word that does not capture the sound, which was the sound of a debt that had been waiting a very long time to be closed. She spoke the name aloud.

The ghost smiled -- the way leaves smile when they release the branch -- and dissolved into the leafmold from which all things in the Valewood are remade. The grove was hers. Truly hers.

\subsubsection{The Scholar's Observation}
\textit{I asked the pathweaver what became of Serevindë. She said: "She became the grove. Not the owner of it -- the grove itself. The trees remember her counting. The western slope has stopped coughing. You can walk there at dusk and hear a woman speaking a name over and over. Not in grief. In practice."}

\textit{I have not walked there. I am told the copper‑beech is very tall.}

\section{Part Four: Through the Mirror of Odd Steps}
\subsection{An Aelinnel Tale of the Green Gate and the Counting of Things That Should Not Be Counted}
\subsubsection{As told by a gnome bell‑tender who refused her name, because names are debts and she has enough. Recorded in the tide‑rift below the Hall of Aelinnel.}
\textit{The Scholar's Note on Methodology: The Aelinnel do not tell stories in the manner of other cultures. They tell them \textit{as equations}. A narrative is a sequence of operations performed on an initial condition (the listener's ignorance) to produce a final condition (the listener's understanding, or sometimes confusion). I have preserved the operations as faithfully as possible. The reader is advised to approach the following with a slate and chalk.}

\textbf{Initial Condition.} A young gnome -- call her \textit{Opal}, though this is not her name; her name is a prime number -- arrived at the Green Gate at the wrong tide. She was counting sheep. The sheep were counting her back. She lost her place.

\textbf{Operation One.} The Gate opened anyway.

Beyond it was not the forest she expected. It was a \textit{room}. The room was a perfect hexagon, each wall a different shade of gray. In the center stood a table laid with a single teacup. The cup was empty. The cup had been empty for a very long time.

\textbf{Operation Two.} A voice spoke from the chair opposite the table. The chair had no occupant. It had a \textit{place} for an occupant -- a dent in the cushion, a warmth on the armrest -- but the person who should have been sitting there was elsewhere, possibly yesterday.

\textit{``You"re early," said the voice. "Tea is served at the seventh hour. It is now the ninth. The ninth is not an hour. It is a \textit{comment}."}

\textbf{Operation Three.} Opal sat. The cup filled itself with tea that was also not tea -- it was the \textit{memory} of tea, which tastes like nostalgia and regret in equal measure.

\textit{``I need to go back," she said.}

\textit{``Everyone does," said the chair. "The question is: \textbf{where?} Back is not a direction. It is a \textbf{consequence}. You can go back to the moment before you counted the sheep, but the sheep have already been counted. Or you can go back to the moment before the sheep were born, but then you would have to count their ancestors, and they are far less cooperative."}

\textbf{Operation Four.} The chair instructed Opal to walk to the darkest gray wall, tap it five times, wait for the echo, and then speak her name -- but only the consonants.

She did.

The wall became a door. The door opened into a corridor that smelled of rain and old mathematics. She walked. She did not count her steps.

\textbf{Final Condition.} At the end of the corridor was the Green Gate, and beyond it, the hawthorn hills of Aelinnel. The tide was right. The Gate was open.

She stepped through.

Behind her, the chair whispered: \textit{``You forgot the tea."}

Opal did not look back.

\subsubsection{The Scholar's Observation}
\textit{I asked the bell‑tender what the chair meant by "a comment." She tapped her temple with a copper nail.}

\textit{``A comment is a number that does not belong in the sequence. It is the sound of a bell that was struck by accident. It is the ninth step on a stair that was measured for eight. It changes nothing -- the sequence continues -- but it cannot be unheard."}

\textit{I asked if the chair was a person. She said: "The chair is a chair. The place is the person. You cannot separate them, but you also cannot confuse them. That is the first lesson of the Green Gate. The second lesson is that the gate does not care if you learn the first."}

\textit{I have not had the opportunity to test this lesson personally. The tide has not been right. I am not certain I want it to be.}

\section{Part Five: The Ninth Step That Became a Door}
\subsection{An Aelaerem Tale of the Hollow, the Threshold, and the Trouble with Skipping}
\subsubsection{As told by an old hedger named Tam, who received a letter from Dusana of the Raven Road and read it aloud in a cider-press barn south of the Moss Road. Recorded by Valerius from Tam's telling, with the letter itself pressed between the pages.}
\textit{Editor's Note: Tam does not read Utaran script. Dusana's letter was written in Aelaerem trade signs, which Tam translated phrase by phrase, sometimes adding his own observations. The following is a synthesis of the letter and Tam's commentary. The original has been deposited with the Synod's archive. It smells of cider and horsehair.}

Cori was twelve, which is young for a halfling and old for a fool. He had chores. He always had chores. His grandmother, who was the Apple-Matron's second cousin and twice as stubborn, believed that idleness invited the Hollow to sit in the empty chairs and watch.

``Count the geese,'' she said. ``HEAL the net. Salt the pork. Sweep the threshold eight times, not nine.''

Cori swept. He counted. One, two, three, four, five, six, seven, eight.

Then he skipped.

Skipping is not a chore. Skipping is the opposite of a chore. His grandmother would have boxed his ears, but she was inside, and the geese were watching, and the geese do not tell tales---they merely remember them, and sometimes honk the punchline at dawn.

He skipped across the lane, over the stile, past the standing stone with the cup-marks. He counted his skips: one, two, three, four, five, six, seven--- ---eight.

His foot landed. The step was the ninth.

He knew it immediately. Not because he had counted wrong---he had counted correctly, that was the problem---but because the world hiccuped. A door appeared in the village wall. It had not been there a moment before. It was made of wood so old it had forgotten how to splinter. Its handle was a finger bone, warm to the touch.

\footnote{The Aelaerem believe that walls, like thresholds, have two sides. The side that faces the village is for the living. The side that faces away is for the Hollow, for the Neighbors, for the debts that have not yet been named. A door that appears unbidden is not an invitation. It is a \textit{reminder} that the wall has always had a door. You simply had not needed it before.}

Cori should have run. He knew the rule: the ninth step is not a step. But he was twelve. He had been sweeping thresholds for three hours. The door was interesting. The door was \textit{new}.

He opened it.

The room inside was not dark. It was not light. It was \textit{preserved}---the light of a cellar that has never been opened, the air of a jar that has held honey for a hundred years and has not yet been asked to let it go.

It was a storage room, or something like one. Shelves lined the walls, carved from a wood that did not belong to any tree Cori knew. The shelves held things:

A bell with no clapper, but a crack in its side that breathed when you leaned close.
A cup that had been poured from, but never emptied---the liquid inside was the color of the sky before a storm, and it did not wet the rim.
A pair of scissors rusted shut, but the rust was fresh, as if they had been cut yesterday.
A candle that had burned down to a stub, but the wick was still warm, and the wax was still soft, and it smelled of grandmother's kitchen and also of the grave.

\footnote{I asked Tam if the objects were magical. He said: ``Everything is magical if you forget to name it. The Aelaerem name their spoons. They do not name this room.''}

Cori walked between the shelves. His footsteps did not echo. The room was too full of \textit{attention} for echoes. He came to the back wall.

Two mirrors faced each other.

They were tall, taller than Cori, framed in dark metal that had been worked into the shape of branches---not carved, but grown, as if the frames had once been living wood and had only recently decided to stop.

The mirrors showed each other. Endless reflections, each smaller than the last, stretching away into a distance that did not exist in any room with walls.

But they did not show Cori.

He saw the back of his own head---the back of it, in the mirror that should have shown his face. The reflection was wrong. He turned. The reflection did not turn with him. It waited.

Between the mirrors, set into the wall, was a plaque of gray stone. The letters were carved in old Aelaerem script---the kind his grandmother used for charms, the kind that hurts your eyes if you look at it too long, because the letters are also knots.

Cori read:

``One mirror is of the Witness. It shows you who you are meant to be, and where you are meant to be. Speak your true name, and you will become that self.

The other mirror is of Aveh. It shows you who you should not be. Speak the name you do not own, and you will be cast out, made anew in a place that does not know you.

Choose wisely. Follow your heart---but your heart is a threshold, and thresholds have teeth.''

\footnote{The phrase ``follow your heart'' appears in several Aelaerem proverbs, always with a qualifier. The full version, which I have since collected, reads: ``Follow your heart, but keep your hand on the door.''}

Cori looked into the left mirror.

He saw himself---but taller, broader, his hair the color of wheat in high summer. He wore leather that smelled of horses, not sheep. Behind him, mountains rose---not the gentle hills of Amedell, but the sharp peaks of the Aelerian range. There was wind on his face, and he was smiling. He could not remember the last time he had smiled like that.

A name formed in his mind. It was not his name. It was a name he had never spoken, never heard, but it fit him like a hand fits a glove that has been waiting for a hand.

He looked into the right mirror.

He saw himself again---but his face was wrong. The eyes were his, but the mouth was someone else's. The clothes were fine, foreign, heavy with embroidery. He was standing in a marble courtyard, and people were bowing, and he was not smiling. He was \textit{pretending} to smile.

No name came from the right mirror. Only a whisper: ``You could be this. You could be worse.''

\footnote{I asked Tam if the right mirror spoke. He said: ``The mirror does not speak. The Hollow does. The mirror is just a wall that happens to be shiny.''}

Cori stepped back. He should not have been there. The door had been a mistake. The ninth step had been a skipping, not a choice.

But the left mirror showed him the wind on the steppe. It showed him a self that was not a halfling, not a hedger's apprentice, not a boy who swept thresholds and counted geese.

He spoke the name.

The wind swept across the Violet Steppe. Through his hair. Through his lungs. He was tall. He was strong. His skin was the color of sun on rock, and his shadow stretched behind him across the grass, and the shadow was \textit{his}---no one else's, not borrowed, not borrowed from a boy who had swept thresholds and counted geese.

The Aelerian mountains loomed to the east, their peaks sharp as a held breath. Behind him, horses plodded and grazed, their bridles jingling. The smell was of dust and sage and something else---something that might have been freedom, or might have been loneliness. It was hard to tell which.

He stood there for a long time.

Then he sat down on a rock and put his head in his hands.

\subsubsection{The Scholar's Observation}
I asked Tam what became of the boy. He said: ``I found him on the steppe, three days' ride from the Moss Road. He was crying. He said he had made a mistake. I brought him home.''

``Did he remember the name?''

Tam was quiet for a long time. Then he said: ``He remembered the name. He forgot which mirror he spoke it into.''

I asked what that meant.

He said: ``It means he does not know if he is the self he was meant to be, or the self he was warned against. He walks the steppe now, sometimes. He visits his grandmother. She pours him cider, but she never fills the cup to the ninth. He counts. He always counts. But he does not skip.''

I asked if the mirrors still exist, waiting in that room behind the door that appears on the ninth step.

Tam stood up. He walked to the threshold of the cider-press barn. He paused, counted to eight under his breath, and stepped out into the dusk.

He did not answer my question.

\subsubsection{The Witch's Closing}
``The ninth step is a door. The door is a mirror. The mirror shows you a self you could be---or a self you already were, before you forgot which side of the wall you belonged on. Choose wisely. Or don't. The door will open again. It always opens again.''

\subsubsection{What Lurks in the Shadow of the Tale}
\textbf{The Unwitnessed Door} --- A threshold that appears only to those who take the ninth step without meaning to. It leads to a room between places, where two mirrors offer two selves. But the room does not guarantee which mirror is which. The Witness and Aveh are not in the habit of labeling their wares.

\textbf{Tags / Mechanics:}
- A character who intentionally takes a ninth step (skipping the eighth, miscounting stairs, pouring the ninth cup) may glimpse the Unwitnessed Door. A Wits + Perception test (DV 4) is required to perceive it; on a failure, the threshold remains invisible, but the character gains the \textbf{Haunted by Possibility} condition---they dream of a door they did not see, and wake with a name on their lips that is not theirs.
- Entering the door requires a second test (Spirit + Resolve, DV 5). On a success, the character reaches the room of mirrors. On a failure, they are pulled into the Hollow for 1d4 hours and emerge with no memory of the experience---but their shadow is ever so slightly taller.
- In the room, the character may choose a mirror. The GM secretly determines which mirror is which (Witness or Aveh) before the choice. The character speaks a name---either their own true name, or the name they believe they are meant to speak.
- \textbf{Outcome} (GM discretion, rolled behind the screen or determined by narrative need):
    - If the mirror was the Witness and the name was true, the character gains a permanent Talent appropriate to their new self (thematic, not purely mechanical---GM and player collaborate).
    - If the mirror was the Witness and the name was false, the character is not transformed, but they \textit{believe} they were. They gain a Condition: \textbf{Uncertain Self} (--1 die to rolls involving identity, memory, or social trust until they resolve the doubt through roleplay).
    - If the mirror was Aveh and the name was spoken, the character is cast out---transported to a location of the GM's choice, possibly far from their home. They gain a new identity (new name, new appearance) but lose one Bond (the people who knew them no longer recognize them).
    - If the mirror was Aveh and the character refused to speak, they are simply returned to the threshold, unharmed but forever wondering.
- \textbf{Clock: The Door's Memory [6]} --- each time the character uses their new self (if transformed), the timer ticks. When full, the door appears again, offering a second choice---or a return to the old self. The cost is a debt: the room will demand something in return for another chance.
- \textbf{Adventure Hook:} An Aelaerem villager has been disappearing into the Unwitnessed Door for years---each time, returning as a different self. A healer, a merchant, a soldier, a ghost. The party must find the door, determine which mirror is which, and help the villager choose a self they can live with---or destroy both mirrors and close the door forever. But the wall may not forgive the closing. And the grandmother still pours cider, waiting for a boy who skipped.

--- Dusana of the Raven Road

\subsubsection{The Scholar's Final, Reluctant Note}
\textit{After Tam finished reading, I asked if he had kept the letter.}

\textit{He reached into his coat and withdrew a folded sheet of vellum, soft as cloth. The ink was faded, but I could make out Dusana's hand---the same looping script I had seen in other ledgers, the same habit of crowding the margins.}

\textit{``She didn't sign it,'' Tam said. ``Not with her name. She drew a raven.''}

\textit{He showed me. The bird was sketched in three lines, its head turned as if looking back at something it had already passed.}

\textit{I asked Tam, at the end of our conversation, whether the boy was happier as a steppe rider or as a hedger's apprentice.}

\textit{Tam said: ``Happiness is not the question. The question is: which debt are you willing to owe? The boy owes himself. He owes the mirror. He owes the wall. He owes the name he spoke and the name he forgot. He will be paying interest for the rest of his life.''}

\textit{``That is not an answer,'' I said.}

\textit{``No,'' said Tam. ``It is a ledger.''}

\textit{He poured me a cup of cider. It was the eighth. He did not pour a ninth.}

\textit{I drank. The cider tasted of apples and iron. I have not skipped since.}

\textbf{Valerius of Thepyrgos}
\textit{Payden's Port, in the fog}
\textit{The raven looks back.}

\section{Conclusion (or, The Inability to Conclude)}
\textit{The Honorable Synod has requested that I provide a summary of findings. I will comply, under protest.}

\textit{The cultures described herein share certain structural features. They count in eights and fear the ninth. They distinguish between debt and theft, between omission and lie, between the presence of a ghost and the presence of a debt that has not yet decided what it wants. They build their architectures -- physical, social, and metaphysical -- on the assumption that the numbers that govern the world are not inherently benevolent, but that they can be negotiated with, if one speaks the correct language and does not ask for more than "enough."}

\textit{Beyond these observations, I have nothing to offer. Every attempt at a universal claim has been contradicted by the next informant, the next footnote, the next foggy night on a broken pier.}

\textit{I have been in the field for two years. I have learned that the ninth bell is not a sound, that the ninth cup is not a drink, that the ninth step is not a step. I have learned that "enough" is not a number but a decision, and that decisions are not subject to the usual rules of arithmetic.}

\textit{I have learned that I do not know how to stop counting.}

\textit{I believe this is the point.}

\textit{Valerius of Thepyrgos}\\
\textit{Payden's Port, in the fog}\\
\textit{The lamp is out. The ink is frozen. The ninth bell --}

\textit{The manuscript breaks off. The remaining pages are blank, except for a single tally‑mark scratched into the margin:}

\textit{I.}

\section*{Colophon}
\textit{This document was written on paper pressed from rag and reed‑mace, in ink cut with lamp‑black and the ash of a burned tariff‑scroll. The binding is gut, tied with eight knots.}

\textit{The ninth knot is not cut.}

\textit{The scholar is still counting.}

\textit{This document contains a letter from Dusana of the Raven Road, copied with permission. The raven remains unsigned.}

\end{document}
