\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{setspace}
\onehalfspacing

\title{The Art of War\\\large Mass Combat, Advanced Tactics, and Strategic Command}
\author{General Marcus Valerius of Ecktoria\\Preparing for the Fourth Sack}
\date{June 30, 2026}

\begin{document}



\begin{quote}
\textit{``The sword is a liar. It promises victory through blood, but blood is the interest on a debt of strategy. The true warrior wins before the blades are drawn—through supply lines, through morale, through the geometry of the battlefield. The clash is merely the payment of a bill already written.''}\\
— General Marcus Valerius of Ecktoria, on the eve of the Fourth Sack
\end{quote}

\section*{Preface: The General's Reflection}
I stand on the western walls of Ecktoria as the first light of dawn touches the Violet Steppe. Below me, the Vilikari cohorts drill with grim precision; beyond them, the Ykrul warbands move like storm clouds on the horizon. This is not my first autumn watching Ecktoria's pride buckle beneath foreign banners—I was a tribune during the Plebian Rebellion, a legate during the Battle at Kennex Falls, and now, as General of the Eastern Marches, I face the inevitable Fourth, and perhaps final sack of Ecktoria.

They call me the "Last Utaran" in the streets, though I wear Ecktoria's crimson and gold with as much pride as any born to the Violet Steppe. Like Aetian before me against Rekedim, I know that victory does not belong to those who shout loudest in the melee, but to those who calculate the weight of grain in a granary, the exit routes from a mountain pass, and the breath between heartbeats in a tense negotiation.

The Vilikari seek not conquest but concession—they want the routes to flow freely, the exits to remain open. The Ykrul warband leader, a stone-cold strategist named Kurgan of the Iron Hoof, understands that Ecktoria's strength lies not in its walls but in its supply chains, its morale, the thousand small decisions that make an army more than a mob. 

This expansion is not for kings in their towers, but for the officers in the mud, the quartermasters counting rations, the scouts reading wind patterns in the grass. It is for those who know that war is won not in the clash of steel, but in the ledger, the latrine, and the last quiet conversation before dawn.

If you read these pages, remember: every arrow spent is a meal not eaten. Every man lost is a field unplowed. War is not a cost to be paid—it is a debt that compounds. Win this battle, and you may lose the next war. 

The road ahead is long. The scars will remain. But if we fight for something worth saving—something that can grow after the fires dim—then those scars will be testimony to what we built, not a monument to what we burned.

\noindent— General Marcus Valerius, Eastern Marches Command, Ecktoria

\input{copyright}
\section{Part I: Foundations of War}
\subsection{The Art of War—A Philosophy}

\subsection{1.1 Core Principles}

The Art of War expansion is built on five principles that extend Fate's Edge's narrative-first philosophy into the realm of armies, campaigns, and large-scale conflict.

\subsubsection{1.1.1 War as Story, Not Simulation}

Mass combat is not a wargame simulation. It is a narrative framework that produces dramatic outcomes. Every army has a story—its origins, its commander, its morale, its supplies. These stories determine outcomes more than any dice roll.

\textbf{Principle:} The battle serves the story, not the other way around.

\subsubsection{1.1.2 Friction and Uncertainty}

\textit{``The plan is the first casualty of contact with the enemy. Trust your training, not your map.''}

War is chaos. The best-laid plans dissolve in smoke, dust, and the terror of combat. This expansion embraces friction—the unpredictability that makes war a crucible of character.

\textbf{Principle:} Plans are starting points, not blueprints. The GM should reward adaptability and punish rigid thinking.

\subsubsection{1.1.3 The Decisive Point}

\textit{``A battle is not a brawl. It is a negotiation with geography, time, and the will of your enemy. The one who finds the decisive point and commits to it wins.''}

Every battle has a center of gravity—the point where the enemy's strength meets their vulnerability. Finding and striking this point is the commander's art.

\textbf{Principle:} Positioning matters more than numbers. A flanking force of fifty can break a regiment of five hundred.

\subsubsection{1.1.4 The Human Element}

\textit{``Soldiers are not numbers. They are names with fears, hopes, and mothers. A commander who forgets this will command an army of ghosts.''}

Morale, exhaustion, loyalty, and fear are the true determinants of battle. A well-fed, rested, and motivated army will defeat a larger but demoralized one.

\textbf{Principle:} Morale is the army's HP. Track it; feed it; protect it.

\subsubsection{1.1.5 The Economy of Force}

\textit{``Every arrow spent is a meal not eaten. Every man lost is a field unplowed. War is not a cost to be paid; it is a debt that compounds.''}

War is a resource game. Supplies, manpower, and time are all currencies. Winning a battle means nothing if it bankrupts your ability to fight the next one.

\textbf{Principle:} Sustainability matters. A pyrrhic victory is a defeat in disguise.

\subsection{1.2 War and the Core Systems}

The Art of War integrates with Fate's Edge's existing mechanics rather than replacing them.

\textbf{Dice Pools:} Unit actions use Attribute + Skill pools, just like individual actions. Commanders roll Leadership + Tactics to issue orders; units roll Melee + Cohesion to attack.

\textbf{Position and Effect:} These apply at every scale. A unit charging downhill is Dominant; a unit caught in a crossfire is Desperate.

\textbf{Story Beats (SB):} The GM uses SB to introduce friction—a messenger killed, a bridge collapsing, a regiment's morale breaking.

\textbf{Boons:} Commanders spend Boons to rally troops, redirect attacks, or gain insight into enemy plans.

\textbf{Harm and Fatigue:} Units have Harm tracks (representing casualties) and Fatigue tracks (representing exhaustion, low supplies, or low morale).

\subsection{1.3 What This Expansion Adds}

\begin{itemize}
  \item \textbf{The War Scale:} A framework for escalating from skirmish to full-scale battle
  \item \textbf{Unit Profiles:} Stat blocks for military formations
  \item \textbf{Command and Control:} Rules for issuing and implementing orders
  \item \textbf{Mass Combat Resolution:} Fast, narrative-driven systems for resolving large battles
  \item \textbf{Logistics and Morale:} The war behind the war
  \item \textbf{Campaign Warfare:} Connecting battles into larger strategic arcs
\end{itemize}

\section{Chapter 2: The War Scale—From Skirmish to Siege}

\subsection{2.1 The Five Scales of War}

War in Fate's Edge can be played at five scales, each with different mechanics and narrative focus.

\subsubsection{2.1.1 Scale I: Skirmish}

\textbf{Scope:} 2–20 combatants on a side\\
\textbf{Resolution:} Standard grid combat (Fate's Edge core rules)\\
\textbf{Focus:} Individual heroics, PC actions, enemy champions\\
\textbf{Duration:} 3–6 rounds

\textbf{When to Use:} Ambushes, raids, duels, small patrol clashes, assassination attempts, rescue missions.

\textbf{Narrative Lens:} The PCs are the decisive factor. Their individual actions can turn the tide.

\subsubsection{2.1.2 Scale II: Raid}

\textbf{Scope:} 20–100 combatants on a side\\
\textbf{Resolution:} Standard combat with unit rules (this expansion)\\
\textbf{Focus:} Unit coordination, commander decisions, flanking\\
\textbf{Duration:} 6–12 rounds

\textbf{When to Use:} Village attacks, caravan defense, fort assaults, river crossings, commando raids.

\textbf{Narrative Lens:} The PCs are unit commanders or specialists. Their decisions shape the engagement.

\subsubsection{2.1.3 Scale III: Battle}

\textbf{Scope:} 100–1,000 combatants on a side\\
\textbf{Resolution:} Mass combat system (this expansion)\\
\textbf{Focus:} Command decisions, morale, terrain, tactical positioning\\
\textbf{Duration:} 12–24 rounds

\textbf{When to Use:} Field battles, fortress sieges, relief operations, major offensives.

\textbf{Narrative Lens:} The PCs are officers, champions, or embedded specialists. The battle is a story with many threads.

\subsubsection{2.1.4 Scale IV: Campaign}

\textbf{Scope:} 1,000–10,000 combatants on a side\\
\textbf{Resolution:} Campaign turn system (this expansion)\\
\textbf{Focus:} Logistics, strategy, diplomacy, supply chains\\
\textbf{Duration:} Weeks to months of in-game time

\textbf{When to Use:} Full-scale wars, invasions, major offensives, wars of attrition.

\textbf{Narrative Lens:} The PCs are commanders, advisors, or special operatives. The war is a tapestry of interlocking events.

\subsubsection{2.1.5 Scale V: War}

\textbf{Scope:} 10,000+ combatants on a side\\
\textbf{Resolution:} Campaign turn system with strategic-level decisions\\
\textbf{Focus:} National strategy, alliances, economy, war aims\\
\textbf{Duration:} Months to years

\textbf{When to Use:} Wars between nations, existential conflicts, apocalyptic struggles.

\textbf{Narrative Lens:} The PCs are generals, diplomats, or key political figures. The fate of nations hangs in the balance.

\subsection{2.2 Choosing the Right Scale}

Ask the table:
\begin{itemize}
  \item What is the story we want to tell about this conflict?
  \item How much detail do we want in resolution?
  \item How much time do we have?
  \item What is the role of the PCs?
\end{itemize}

\textbf{Quick Reference:}

\begin{tabular}{|l|l|}
\hline
Desired Experience & Recommended Scale \\
\hline
``We're heroes'' & Scale I (Skirmish) \\
\hline
``We're commanders'' & Scale II (Raid) \\
\hline
``We're in a war'' & Scale III (Battle) \\
\hline
``We're shaping history'' & Scale IV–V (Campaign/War) \\
\hline
\end{tabular}

\subsection{2.3 Escalation and De-escalation}

A campaign can move between scales fluidly.

\textbf{Escalating:} A skirmish becomes a battle when reinforcements arrive. A battle becomes a campaign when it becomes a series of engagements.

\textbf{De-escalating:} A battle reduces to a skirmish when the main forces withdraw, leaving only the PCs and a handful of enemies to resolve the final confrontation.

\textbf{Mechanical Transition:} When the scale changes, reset Position and adjust timers. The GM should narrate the transition clearly.

\section{Part II: Advanced Tactics}

\section{Chapter 3: The Tactical Layer—Command and Control}

\subsection{3.1 The Command Structure}

Every military force has a command hierarchy. Understanding this hierarchy helps players make meaningful decisions.

\subsubsection{3.1.1 Command Roles}

\textbf{Commander-in-Chief (Strategeōs)}\\
\begin{itemize}
  \item Sets strategic objectives
  \item Decides when and where to engage
  \item Allocates forces between fronts
  \item \textbf{Mechanics:} Makes command rolls for major decisions; can spend Boons to inspire or redirect
\end{itemize}

\textbf{Adjutant/Staff Officer}\\
\begin{itemize}
  \item Assists the commander with communications, intelligence, and logistics
  \item Coordinates between units and manages information flow
  \item \textbf{Mechanics:} Can reroll a failed Command roll once per battle; grants +1 Position to all units when succeeding on a Wits + Tactics roll to interpret enemy movements
\end{itemize}

\textbf{Battlefield Commander (Tactikos)}\\
\begin{itemize}
  \item Commands forces in a specific engagement
  \item Makes tactical decisions (flanking, reserves, feints)
  \item \textbf{Mechanics:} Issues orders to unit leaders; makes rolls for tactical maneuors
\end{itemize}

\textbf{Unit Commander (Lochagos)}\\
\begin{itemize}
  \item Commands a single unit (company, squadron, war-band)
  \item Leads from the front
  \item \textbf{Mechanics:} Directs unit's actions; personal combat can inspire or demoralize
\end{itemize}

\textbf{Specialist/Champion}\\
\begin{itemize}
  \item PC with a specific role (healer, mage, scout, champion)
  \item Acts independently within the larger engagement
  \item \textbf{Mechanics:} Personal actions can affect unit morale, enemy strategy, or key objectives
\end{itemize}

\subsubsection{3.1.2 Chain of Command}

In most military forces, orders flow from the top down. Disobeying orders has consequences.

\textbf{Mechanical Effect:} When a commander issues an order, units that obey gain +1 Position. Units that disobey suffer -1 Position (or face disciplinary action after the battle).

\textbf{Breaking the Chain:} PCs can disobey orders for dramatic effect. This should cost something—a Boon, a ticking Morale timer, or a relationship with the commander.

\subsection{3.2 Issuing Orders}

Orders are issued during a Command Phase that occurs before the Engagement Phase.

\subsubsection{3.2.1 Command Phase Procedure}

\begin{enumerate}
  \item Assess Situation: The commander evaluates the current state of the battle.
  \item Declare Intent: The commander states what they want each unit to do.
  \item Make a Command Roll: Roll Presence + Tactics (or appropriate skill).
  \item Apply Results: Based on the outcome, units may gain bonuses or penalties to their actions.
\end{enumerate}

\subsubsection{3.2.2 Command Roll Outcomes}

\begin{tabular}{|l|l|}
\hline
Outcome & Effect \\
\hline
Critical Success & All units gain +1 Position this round; commander gains 1 Boon \\
\hline
Success & All units act normally; one unit may act at +1 Position \\
\hline
Partial & One unit acts at +1 Position, one at -1 Position \\
\hline
Miss & All units act at -1 Position; commander may spend 1 Boon to negate \\
\hline
\end{tabular}

\subsubsection{3.2.3 Types of Orders}

\textbf{Simple Orders:} ``Advance,'' ``Hold,'' ``Fall Back.''\\
\begin{itemize}
  \item No roll required to understand
  \item All units can obey
\end{itemize}

\textbf{Complex Orders:} ``Flanking Maneuver,'' ``Feigned Retreat,'' ``Envelopment.''\\
\begin{itemize}
  \item Requires a Tactics roll from the unit commander
  \item May take one full round to set up
\end{itemize}

\textbf{Contingency Orders:} ``If the enemy breaks, pursue; if they hold, withdraw.''\\
\begin{itemize}
  \item A clever commander anticipates events
  \item Units with contingency orders act as if commanded when the trigger occurs
\end{itemize}

\subsection{3.3 Initiative and Turn Order}

At the tactical level, combat follows an engagement sequence.

\subsubsection{3.3.1 Engagement Sequence}

\begin{enumerate}
  \item Command Phase: All commanders issue orders.
  \item Movement Phase: Units move according to orders.
  \item Ranged Phase: Ranged units fire.
  \item Melee Phase: Melee units engage.
  \item Morale Phase: Units check morale based on casualties and events.
  \item End Phase: Resolve ongoing effects; reset for next turn.
\end{enumerate}

\subsubsection{3.3.2 Initiative}

Initiative at the tactical level is determined by:
\begin{itemize}
  \item Commander's Initiative: Add Wits + Tactics to determine order.
  \item Unit Speed: Faster units act sooner in the Movement Phase.
  \item Surprise: Ambushing units act first; surprised units are Desperate.
\end{itemize}

\textbf{Mechanical Effect:} The side with initiative moves first in each phase. If initiative is tied, both sides act simultaneously.

\subsection{3.4 Communications and the Fog of War}

\textit{``The first casualty of battle is a clear picture of what is happening. Trust your eyes, not your intelligence.''}

War is inherently chaotic. Communications fail, misinformation spreads, and the battlefield is often obscured.

\subsubsection{3.4.1 Communication Failure}

When a unit is separated from its commander, orders may not reach them.

\textbf{Causes:}
\begin{itemize}
  \item Courier killed (GM spends 1 SB)
  \item Smoke, dust, or fog obscuring signals
  \item Enemy jamming or intercepting signals
  \item Unit moves out of communication range
\end{itemize}

\textbf{Mechanical Effect:} A unit without orders acts on its own initiative (GM decides or makes a simple roll). The unit may also suffer -1 Position due to confusion.

\textbf{Prevention:}
\begin{itemize}
  \item Use multiple couriers
  \item Establish pre-arranged signals (flags, horns, etc.)
  \item Keep units close enough for direct communication
\end{itemize}

\subsubsection{3.4.2 The Fog of War Table (d6)}

When the GM spends 2 SB, roll on this table for a ``fog of war'' complication:

\begin{tabular}{|l|l|}
\hline
d6 & Complication \\
\hline
1 & False Report: A unit is reported dead or routing when they are not. \\
\hline
2 & Misinterpreted Orders: A unit takes the wrong action (attack instead of hold, etc.). \\
\hline
3 & Delayed Courier: Orders arrive one round late. \\
\hline
4 & Signal Error: A signal is misread; a flanking force attacks a friendly unit. \\
\hline
5 & Rumor Spreads: A rumor of defeat demoralizes a unit (Morale -1). \\
\hline
6 & Confusion Reigns: All units suffer -1 Position this round as the fog of war thickens. \\
\hline
\end{tabular}

\section{Chapter 4: Unit Profiles and Capabilities}

\subsection{4.1 Anatomy of a Unit}

A unit is a collection of soldiers represented by a stat block. Units are the basic building blocks of tactical combat.

\subsubsection{4.1.1 Unit Stat Block}

\textbf{Unit Name:} The unit's designation\\
\textbf{Type:} Heavy Infantry, Light Cavalry, Archers, etc.\\
\textbf{Capability (Cap):} 1–5 (see below)\\
\textbf{Scale:} I (Squad), II (Company), III (Battalion), IV (Regiment)\\
\textbf{Speed:} Base movement rate (varies by type)\\
\textbf{Tags:} Special qualities (Elite, Siege, Skirmish, etc.)\\

\textbf{Tracks:}
\begin{itemize}
  \item Cohesion (6): Morale and discipline
  \item Exposure (6): Supplies, fatigue, and exhaustion
  \item Resources (6): Ammunition, equipment, and rations
\end{itemize}

\textbf{Orders:} What this unit can do (see Chapter 5)

\subsubsection{4.1.2 Capability (Cap)}

Capability represents the unit's training, experience, and equipment.

\begin{tabular}{|l|l|l|}
\hline
Cap & Description & Example \\
\hline
1 & Levy/Recruits & Peasant levy, raw recruits \\
\hline
2 & Trained & Standard infantry, militia \\
\hline
3 & Veteran & Professional soldiers, experienced mercenaries \\
\hline
4 & Elite & Guard troops, elite mercenaries \\
\hline
5 & Legendary & Praetorians, champions, legendary heroes \\
\hline
\end{tabular}

\textbf{Cap Determines:}
\begin{itemize}
  \item Die Pool: Cap + Skill when acting
  \item Morale: Higher Cap units resist fear and break less often
  \item Damage: Higher Cap units deal more Harm
\end{itemize}

\textbf{Cap and Tier Relation:}
\begin{tabular}{|l|l|}
\hline
Tier & Typical Unit Cap \\
\hline
Tier I (Novice) & Cap 1-2 (Levy/Trained) \\
\hline
Tier II (Seasoned) & Cap 2-3 (Trained/Veteran) \\
\hline
Tier III (Veteran) & Cap 3-4 (Veteran/Elite) \\
\hline
Tier IV (Paragon) & Cap 4 (Elite) \\
\hline
Tier V (Mythic) & Cap 5 (Legendary) \\
\hline
\end{tabular}

\subsubsection{4.1.3 Scale}

Scale represents the size of the unit.

\begin{tabular}{|l|l|l|}
\hline
Scale & Size & Examples \\
\hline
I & 1–10 & Squad, warband \\
\hline
II & 10–50 & Company, cavalry squadron \\
\hline
III & 50–200 & Battalion, war-band \\
\hline
IV & 200–1,000 & Regiment, army \\
\hline
\end{tabular}

\textbf{Scale Determines:}
\begin{itemize}
  \item Area of Effect: Larger units cover more ground
  \item ZoC: Larger units project stronger zones of control
  \item Targetability: Larger units are easier to hit with area effects
\end{itemize}

\subsection{4.2 Unit Types}

Different unit types have different strengths, weaknesses, and tactical roles.

\subsubsection{4.2.1 Infantry}

\textbf{Role:} Hold the line, engage in melee, occupy key positions

\textbf{Heavy Infantry (Cap 3–4)}\\
\begin{itemize}
  \item Speed: Slow
  \item Defense: High (armor, shields)
  \item Offense: High (2-handed weapons)
  \item Special: Can ``Set'' against cavalry
  \item Weakness: Poor in rough terrain; vulnerable to skirmishers
\end{itemize}

\textbf{Medium Infantry (Cap 2–3)}\\
\begin{itemize}
  \item Speed: Standard
  \item Defense: Moderate (shields)
  \item Offense: Moderate (sword and shield)
  \item Special: Versatile, can hold or attack
  \item Weakness: Not exceptional at anything
\end{itemize}

\textbf{Light Infantry (Cap 2–3)}\\
\begin{itemize}
  \item Speed: Fast
  \item Defense: Low (light armor)
  \item Offense: Low (javelins, short swords)
  \item Special: Skirmishing, can harass enemy flanks
  \item Weakness: Cannot hold the line
\end{itemize}

\subsubsection{4.2.2 Cavalry}

\textbf{Role:} Flank, charge, pursue, disrupt enemy formations

\textbf{Heavy Cavalry (Cap 3–4)}\\
\begin{itemize}
  \item Speed: Fast
  \item Defense: High (armor, shields)
  \item Offense: High (lances)
  \item Special: Devastating charge; breakthrough
  \item Weakness: Poor in rough terrain; vulnerable to prepared infantry
\end{itemize}

\textbf{Medium Cavalry (Cap 2–3)}\\
\begin{itemize}
  \item Speed: Fast
  \item Defense: Moderate
  \item Offense: Moderate
  \item Special: Flexible; can skirmish or charge
  \item Weakness: Not as powerful as heavy cavalry
\end{itemize}

\textbf{Light Cavalry (Cap 1–3)}\\
\begin{itemize}
  \item Speed: Very fast
  \item Defense: Low
  \item Offense: Low (bows, javelins)
  \item Special: Skirmishing, reconnaissance
  \item Weakness: Cannot hold against infantry
\end{itemize}

\subsubsection{4.2.3 Skirmishers}

\textbf{Role:} Harass, screen, scout, weaken enemy formations

\textbf{Archers (Cap 2–3)}\\
\begin{itemize}
  \item Speed: Standard
  \item Defense: Low
  \item Offense: Ranged
  \item Special: High rate of fire; suppressive fire
  \item Weakness: Poor in melee
\end{itemize}

\textbf{Crossbowmen (Cap 2–3)}\\
\begin{itemize}
  \item Speed: Slow
  \item Defense: Low
  \item Offense: Ranged, armor-piercing
  \item Special: Armor penetration; devastating volleys
  \item Weakness: Slow to reload
\end{itemize}

\textbf{Skirmishers/Javelinmen (Cap 2)}\\
\begin{itemize}
  \item Speed: Fast
  \item Defense: Low
  \item Offense: Ranged, short range
  \item Special: Harass; can engage and disengage
  \item Weakness: Cannot hold; short range
\end{itemize}

\subsubsection{4.2.4 Special Units}

\textbf{Engineers (Cap 2–3)}\\
\begin{itemize}
  \item Role: Siege works, fortifications, bridges, demolition
  \item Special: Can construct or demolish terrain features
  \item Weakness: Not combat units; need protection
\end{itemize}

\textbf{Artillery (Cap 2–3)}\\
\begin{itemize}
  \item Role: Bombard, siege, area denial
  \item Special: Area attacks; devastating against formations
  \item Weakness: Slow to set up; vulnerable to flanking
\end{itemize}

\textbf{Magical Support (Cap 2–4)}\\
\begin{itemize}
  \item Role: Buff, debuff, area damage, communication
  \item Special: Can alter the battlefield in unexpected ways
  \item Weakness: Magical energy is finite; vulnerable to counterspelling
\end{itemize}

\subsection{4.3 Unit Tags}

Tags are special qualities that modify a unit's behavior.

\begin{tabular}{|l|l|}
\hline
Tag & Effect \\
\hline
Elite & +1 Cap; better morale \\
\hline
Veteran & Ignore first Morale check per battle \\
\hline
Green & -1 Cap; worse morale \\
\hline
Siege & +1 Effect vs. fortifications; slow movement \\
\hline
Skirmish & Can engage and disengage at +1 Position \\
\hline
Pikes & +1 Effect vs. cavalry \\
\hline
Shield Wall & +1 Defense when stationary \\
\hline
Raiders & +1 Speed; better loot \\
\hline
Fanatic & -1 Morale, but won't rout (dies fighting) \\
\hline
Mercenary & Fights for pay; may switch sides \\
\hline
Magic & Can use magical effects; vulnerable to counterspelling \\
\hline
\end{tabular}

\subsection{4.4 Follower Units}

Use the Follower system (see Player's Guide, Chapter 12) for unit commanders and embedded specialists.

\subsubsection{4.4.1 Unit Commander Template}

\textbf{Name:} [Unit Name]\\
\textbf{Type:} [Unit Type]\\
\textbf{Cap:} 2–4\\
\textbf{Scale:} I–III\\
\textbf{Tags:} [As appropriate]\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6]\\
\textbf{Orders:} [As appropriate]

\subsubsection{4.4.2 Embedded Specialist Template}

\textbf{Name:} [PC or NPC Name]\\
\textbf{Role:} Champion, mage, healer, scout\\
\textbf{Cap:} 1–5 (individual capability)\\
\textbf{Action:} Acts independently within the larger engagement

\section{Chapter 5: Maneuvers, Formations, and Special Actions}

\subsection{5.1 Basic Maneuvers}

Units can perform a variety of maneuvers, each with distinct effects.

\subsubsection{5.1.1 Advance}

The unit moves forward in an orderly fashion.

\textbf{Effect:} Move standard speed; maintain formation.\\
\textbf{Position:} Controlled (if no enemy contact) or Standard (if engaged)\\
\textbf{Risk:} Vulnerable during movement.

\subsubsection{5.1.2 Charge}

The unit moves forward with the intention of engaging violently.

\textbf{Effect:} Move speed + 2; gain +1 Effect on first attack.\\
\textbf{Position:} Dominant (if from high ground or flank)\\
\textbf{Risk:} -1 Defense until next turn; vulnerable to prepared enemies.

\textbf{Special:} Cavalry charge has +2 Effect instead of +1.

\subsubsection{5.1.3 Withdraw}

The unit falls back in an orderly fashion.

\textbf{Effect:} Move standard speed; maintain formation.\\
\textbf{Position:} Controlled\\
\textbf{Risk:} May be pursued; +1 Morale check if pursued.

\textbf{Special:} If the unit is engaged, the Withdraw requires a successful Morale check.

\subsubsection{5.1.4 Rout}

The unit flees in disorder.

\textbf{Effect:} Move speed + 2; drop all formation bonuses.\\
\textbf{Position:} Desperate\\
\textbf{Risk:} Enemies gain +1 Position on all attacks against routing units.

\textbf{Special:} Routing units can be rallied by a commander (see Rally).

\subsection{5.2 Formations}

Formations are special arrangements that provide bonuses and create vulnerabilities.

\subsubsection{5.2.1 Shield Wall}

\textbf{Requirements:} Shields\\
\textbf{Bonus:} +2 Defense, +1 Morale\\
\textbf{Penalty:} -1 Speed, cannot charge\\
\textbf{Vulnerability:} Flanking, artillery, magical attacks

\textbf{Mechanical Effect:} The unit is a solid block. Enemies attacking from the front suffer -1 Position.

\subsubsection{5.2.2 Schiltron}

\textbf{Requirements:} Spears/Pikes\\
\textbf{Bonus:} +2 vs. Cavalry, +1 Defense\\
\textbf{Penalty:} -1 Speed, rigid formation\\
\textbf{Vulnerability:} Flanking, long-range, magic

\textbf{Mechanical Effect:} Cavalry charges against this unit suffer -2 Effect; the unit is resistant to breakthrough.

\subsubsection{5.2.3 Wedge (Cuneus)}

\textbf{Requirements:} Heavy infantry or cavalry\\
\textbf{Bonus:} +2 Charge, +1 Breakthrough\\
\textbf{Penalty:} Vulnerable to flanking, -1 Defense\\
\textbf{Vulnerability:} Flanking, Artillery

\textbf{Mechanical Effect:} The unit punches through enemy lines with +1 Position on the first round of combat.

\subsubsection{5.2.4 Skirmish Line}

\textbf{Requirements:} Skirmishers\\
\textbf{Bonus:} +1 Movement, +1 Ranged\\
\textbf{Penalty:} Cannot hold, -1 Defense in melee\\
\textbf{Vulnerability:} Heavy units, cavalry

\textbf{Mechanical Effect:} The unit screens other units; can engage and disengage freely.

\subsubsection{5.2.5 Column}

\textbf{Requirements:} Any\\
\textbf{Bonus:} +2 Speed on roads\\
\textbf{Penalty:} -1 Defense, -1 Melee\\
\textbf{Vulnerability:} Ambush, artillery

\textbf{Mechanical Effect:} The unit moves quickly but is vulnerable to attack.

\subsection{5.3 Special Actions}

\subsubsection{5.3.1 Rally}

\textbf{Commander Action:} Attempt to restore a unit's Morale or Cohesion.

\textbf{Roll:} Presence + Command vs. DV (unit's current Cohesion deficit)\\
\textbf{Effect:} Restore up to 2 Cohesion on success; partial restores 1.

\subsubsection{5.3.2 Flank}

\textbf{Unit Action:} Attempt to attack from the side or rear.

\textbf{Roll:} Tactics vs. enemy's Command (opposed roll)\\
\textbf{Effect:} Success gives the unit +1 Position on its next attack. A critical success gives +2 Position (Dominant).

\subsubsection{5.3.3 Feint}

\textbf{Unit Action:} Deceive the enemy about true intentions.

\textbf{Roll:} Deception vs. enemy's Perception (opposed roll)\\
\textbf{Effect:} Success makes the enemy react in a predictable way; the next friendly action gains +1 Position. Partial creates confusion (both sides -1 Position).

\subsubsection{5.3.4 Set Against Charge}

\textbf{Unit Action:} Prepare to receive a charge.

\textbf{Roll:} Tactics + Command (DV 3)\\
\textbf{Effect:} Success gives +2 Effect vs. charging enemies this round. Failure leaves the unit vulnerable (Desperate vs. charges).

\subsubsection{5.3.5 Volley Fire}

\textbf{Ranged Unit Action:} Concentrated arrow fire on a target area.

\textbf{Effect:} Area attack; all units in the target zone suffer Harm 1. On a critical success, Harm 2. This action costs 1 Resource (ammunition).

\subsubsection{5.3.6 Suppressive Fire}

\textbf{Ranged Unit Action:} Pin down enemy units with constant fire.

\textbf{Effect:} Affected units suffer -1 Position until they move out of the suppression zone. Success suppresses for one round; critical success for two.

\subsubsection{5.3.7 Breach}

\textbf{Infantry Action:} Attempt to break through an enemy line.

\textbf{Roll:} Melee + Cohesion vs. enemy's Melee + Cohesion (opposed)\\
\textbf{Effect:} Success opens a gap in the enemy line; units can pass through. Failure causes the unit to be repulsed (Morale -1).

\subsubsection{5.3.8 Hold}

\textbf{Infantry Action:} Refuse to give ground.

\textbf{Effect:} The unit stays in position. If attacked, it gains +1 Defense. If it takes casualties, it must make a Morale check to remain.

\section{Chapter 6: Terrain and Environment}

\subsection{6.1 Terrain Types}

Terrain affects movement, cover, visibility, and combat effectiveness.

\subsubsection{6.1.1 Open Field}

\textbf{Movement:} Normal\\
\textbf{Cover:} None (or limited)\\
\textbf{Visibility:} Good\\
\textbf{Bonuses:} Cavalry charge, artillery, mass formations\\
\textbf{Penalties:} Vulnerable to ranged attacks, no cover

\subsubsection{6.1.2 Hill}

\textbf{Movement:} Up = 1.5x, down = normal\\
\textbf{Cover:} Partial (on ridge)\\
\textbf{Visibility:} Good from top, limited from bottom\\
\textbf{Bonuses:} +1 Position for defenders on high ground\\
\textbf{Penalties:} Slower movement uphill

\subsubsection{6.1.3 Forest}

\textbf{Movement:} 0.5x\\
\textbf{Cover:} Good (trees)\\
\textbf{Visibility:} Poor (about 2–3 cells)\\
\textbf{Bonuses:} Ambush; skirmishers; concealment\\
\textbf{Penalties:} No cavalry charges; reduced missile range; -1 Command

\subsubsection{6.1.4 Marsh}

\textbf{Movement:} 0.25x or impossible\\
\textbf{Cover:} None\\
\textbf{Visibility:} Poor (fog)\\
\textbf{Bonuses:} Defensive position; slows enemy\\
\textbf{Penalties:} Units are vulnerable to fire; -1 Melee; heavy units become Desperate

\subsubsection{6.1.5 Urban/Ruin}

\textbf{Movement:} 0.5x (streets), variable (buildings)\\
\textbf{Cover:} Good (buildings)\\
\textbf{Visibility:} Limited (line of sight around corners)\\
\textbf{Bonuses:} Defensive advantage; ambush; small unit tactics\\
\textbf{Penalties:} Large units suffer -1 Position; no cavalry; command and control difficult

\subsubsection{6.1.6 Water}

\textbf{Movement:} 0.5x (wading), requires boats (deep)\\
\textbf{Cover:} None\\
\textbf{Visibility:} Poor (mists, reflections)\\
\textbf{Bonuses:} Defensive barrier; slows enemy\\
\textbf{Penalties:} Vulnerable to attack while crossing; heavy units are Desperate

\subsection{6.2 Terrain Mechanics}

\subsubsection{6.2.1 Cover}

\begin{tabular}{|l|l|l|}
\hline
Cover Type & Defense Bonus & Effect \\
\hline
None & 0 & Normal \\
\hline
Light & +1 & Trees, low walls, partial cover \\
\hline
Medium & +2 & Stone walls, fortifications \\
\hline
Heavy & +3 & Fortress walls, bunkers \\
\hline
Total & Immune (until breached) & Castle walls, impenetrable \\
\hline
\end{tabular}

\textbf{Mechanical Effect:} Cover gives +1 Defense per level; can be destroyed by siege engines or magic (GM spends SB to degrade cover).

\subsubsection{6.2.2 High Ground}

\textbf{Mechanical Effect:} A unit on higher ground gains +1 Position when attacking downhill and +1 Defense when defending from below.

\subsubsection{6.2.3 Obscured Terrain}

\textbf{Mechanical Effect:} When line of sight is obscured, missile attacks are at -1 die. Ambushes are easier (Deception +1).

\subsection{6.3 Weather}

Weather can dramatically affect a battle.

\subsubsection{6.3.1 Weather Conditions Table (d6)}

\begin{tabular}{|l|l|l|}
\hline
d6 & Condition & Effect \\
\hline
1 & Clear & Normal \\
\hline
2 & Light Rain & -1 missile range, -1 visibility; Morale -1 \\
\hline
3 & Heavy Rain & -2 missile range, -2 visibility; movement 0.5x; Morale -1 \\
\hline
4 & Snow & -2 movement; -1 visibility; cold effects (Fatigue from exposure); Morale -1 \\
\hline
5 & Fog & Visibility 3 cells; -2 missile range; ambushes easier; Morale -1 \\
\hline
6 & Storm & No missile fire; movement 0.25x; units may rout from fear; lightning can strike; Morale -2 \\
\hline
\end{tabular}

\subsubsection{6.3.2 Weather Changes}

The GM may change weather conditions during a battle by spending 2 SB or at the end of a round.

\textbf{Mechanical Effect:} Weather change gives one side or the other a temporary advantage or disadvantage. The GM should narrate the change dramatically.

\subsection{6.4 Terrain and Cover: Quick Reference}

\begin{tabular}{|l|l|}
\hline
Situation & Effect \\
\hline
Light Cover & +1 Defense vs. ranged attacks \\
\hline
Medium Cover & +2 Defense vs. ranged attacks; +1 Defense vs. melee \\
\hline
Heavy Cover & +3 Defense vs. ranged attacks; +2 Defense vs. melee \\
\hline
High Ground & +1 Position attacking downhill; +1 Defense defending \\
\hline
Difficult Terrain & Movement 0.5x; -1 Position to heavy units \\
\hline
Rough Terrain & Movement 0.25x; -1 Position to all units \\
\hline
Visibility Poor & Missile attacks -1 die; ambushes easier \\
\hline
\end{tabular}

\section{Part III: Mass Combat}

\section{Chapter 7: Bridging the Gap—From Squad to Army}

\subsection{7.1 A Word from the General}

\textit{``I have stood on the hill overlooking a thousand men, knowing that each of them had a mother who would weep for them—and a captain who might die if I made the wrong choice. The weight of that moment is not mechanical. It is moral. If your mass combat system does not make you feel that weight, you are not using it right.''}

\subsection{7.2 Design Philosophy}

Mass combat in Fate's Edge should be:
\begin{enumerate}
  \item Fast: Not slower than regular combat
  \item Narrative: Focused on decisions, drama, and consequences
  \item Heroic: PCs are still the main characters
  \item Meaningful: The outcome matters
\end{enumerate}

\subsection{7.3 The Mass Combat System}

The mass combat system is a streamlined way to resolve battles involving hundreds or thousands of troops. It focuses on the decisions that commanders make and the moments of heroism that change battles.

\subsubsection{7.3.1 Components}

\begin{itemize}
  \item Battle Map: Terrain and positions
  \item Units: Each with a stat block
  \item Commanders: PCs and key NPCs
  \item Objectives: What each side wants
  \item Timers: Deadlines, reinforcements, key events
\end{itemize}

\subsubsection{7.3.2 Battle Phases}

\begin{enumerate}
  \item Planning Phase: Players and GM decide objectives, unit positions, and strategies.
  \item Command Phase: Orders are issued.
  \item Engagement Phase: Units act and interact.
  \item Resolution Phase: Casualties, morale, and objective progress are resolved.
  \item End Phase: The battle is concluded.
\end{enumerate}

\subsection{7.4 Unit Actions in Mass Combat}

Units in mass combat have three action types.

\subsubsection{7.4.1 Standard Actions}

\begin{itemize}
  \item Advance: Move and attack (or prepare to attack)
  \item Hold: Defend current position
  \item Withdraw: Fall back in good order
\end{itemize}

\subsubsection{7.4.2 Complex Actions}

\begin{itemize}
  \item Flank: Attempt to attack from side or rear
  \item Charge: Attack with momentum
  \item Feint: Trick the enemy
  \item Volley: Concentration of missile fire
\end{itemize}

\subsubsection{7.4.3 Special Actions}

\begin{itemize}
  \item Rally: Boost morale
  \item Breach: Break enemy line
  \item Screen: Protect other units
  \item Set: Prepare for enemy action
\end{itemize}

\subsection{7.5 The Mass Combat Roll}

Mass combat uses an opposed roll system:

\textbf{Unit Roll:} Cap + Skill (Melee, Ranged, etc.) + Boons - Penalties\\
\textbf{Opposed Roll:} Compare results. Higher roll wins.\\

\textbf{Effect:}
\begin{itemize}
  \item Success: Enemy takes casualties; objective advances
  \item Partial: Stalemate; both sides take casualties
  \item Miss: Unit suffers casualties; objective recedes
\end{itemize}

\subsection{7.6 The Area of Engagement}

The battlefield is divided into zones of control.

\textbf{Zones:}
\begin{itemize}
  \item Frontline: Where units are actively engaged
  \item Reserve: Behind the frontline; ready to reinforce
  \item Flank: Side areas; vulnerable to attack
  \item Rear: Behind the army; supplies and command
\end{itemize}

\textbf{Mechanical Effect:} Moving between zones costs an action. Flanking attacks gain +1 Position.

\section{Chapter 8: Mass Combat Resolution}

\subsection{8.1 The Battle Turn}

A battle turn represents a period of intensive combat (usually 10–30 minutes).

\subsubsection{8.1.1 Turn Sequence}

\begin{enumerate}
  \item Command Phase
    \begin{itemize}
      \item Commanders issue orders to units
      \item Command rolls are made (see Chapter 3)
      \item Objectives are updated
    \end{itemize}
  \item Movement Phase
    \begin{itemize}
      \item Units move according to orders
      \item Movement is resolved using movement rates and terrain rules
    \end{itemize}
  \item Ranged Phase
    \begin{itemize}
      \item Missile units fire
      \item Artillery fires (if present)
      \item Area attacks are resolved
    \end{itemize}
  \item Melee Phase
    \begin{itemize}
      \item Engaged units fight
      \item Special actions (charge, breach, etc.) are resolved
    \end{itemize}
  \item Morale Phase
    \begin{itemize}
      \item Units check morale
      \item Casualties are tallied
      \item Unit statuses are updated
    \end{itemize}
  \item End Phase
    \begin{itemize}
      \item Commanders may issue new orders for next turn
      \item Special events are resolved (reinforcements, supply issues, etc.)
    \end{itemize}
\end{enumerate}

\subsection{8.2 Resolving Combat}

\subsubsection{8.2.1 Attack Roll}

\textbf{Base Roll:} Unit Cap + Relevant Skill\\
\textbf{Examples:}
\begin{itemize}
  \item Melee unit: Cap + Melee
  \item Ranged unit: Cap + Ranged
  \item Magic unit: Cap + Arcana
\end{itemize}

\textbf{Modifiers:}
\begin{itemize}
  \item Position: +1 for Dominant, -1 for Desperate
  \item Terrain: +1 for favorable, -1 for unfavorable
  \item Morale: +1 for high morale, -1 for low morale
  \item Fatigue: -1 for moderate fatigue, -2 for severe fatigue
\end{itemize}

\textbf{Result:} Compare to enemy's Defense Value (DV). Success deals Harm.

\subsubsection{8.2.2 Unit Harm}

Units have Harm tracks based on Cap.

\begin{tabular}{|l|l|}
\hline
Cap & Harm Levels \\
\hline
1 & 2 \\
\hline
2 & 3 \\
\hline
3 & 4 \\
\hline
4 & 5 \\
\hline
5 & 6 \\
\hline
\end{tabular}

\textbf{Harm Effects:}
\begin{itemize}
  \item 1 Harm: -1 die to all rolls
  \item 2 Harm: -2 dice to all rolls
  \item 3 Harm: -3 dice to all rolls (unit is crippled)
  \item 4+ Harm: Unit is destroyed or routed
\end{itemize}

\subsubsection{8.2.3 Morale}

Morale is tracked as Cohesion (6). It is affected by:

\textbf{Morale Bonuses:}
\begin{itemize}
  \item High ground: +1
  \item Command success: +1 (temporary)
  \item Elite unit: +1 (permanent)
  \item Champion presence: +1
\end{itemize}

\textbf{Morale Penalties:}
\begin{itemize}
  \item Casualties: -1 per level of Harm taken
  \item Friendly unit routing: -1
  \item Commander killed: -2
  \item Supply shortage: -1
\end{itemize}

\textbf{Morale Check:}
\begin{itemize}
  \item Roll Cohesion + Command vs. DV (current Cohesion deficit)
  \item Success: Unit holds
  \item Partial: Unit is shaken (-1 Position next turn)
  \item Miss: Unit routs
\end{itemize}

\subsection{8.3 Routing}

When a unit's Cohesion reaches 0 or it fails a Morale check, it routs.

\textbf{Rout Effects:}
\begin{itemize}
  \item Unit moves at double speed away from the enemy
  \item Unit drops all formation bonuses
  \item Unit is Desperate Position for all purposes
  \item Other units suffer -1 Morale (friendly units may rout in sympathy)
\end{itemize}

\textbf{Rallying:}
\begin{itemize}
  \item Commander Action: Rally
  \item Roll Presence + Command vs. DV (unit's current Cohesion deficit)
  \item Success: Unit stops routing and reforms (Cohesion 1)
  \item Critical Success: Unit returns to the fight with Cohesion 2
\end{itemize}

\subsection{8.4 The Cost of Victory}

Mass combat should feel costly. Losing should be devastating. Winning should feel hard-won.

\textbf{Casualty Tally:}
\begin{itemize}
  \item Each unit that routs or is destroyed suffers permanent casualties
  \item Casualties are tracked as ``Lost Cohesion'' that may take weeks to recover
  \item Hired units (mercenaries) who suffer heavy casualties may demand higher pay or leave
\end{itemize}

\textbf{Aftermath:}
\begin{itemize}
  \item Units need time to recover
  \item Supplies must be replenished
  \item Morale must be rebuilt
  \item Dead must be buried
\end{itemize}

\section{Chapter 9: Commanders, Heroes, and the Friction of War}

\subsection{9.1 The Commander's Role}

A commander does not fight. A commander decides. Their role is to:
\begin{enumerate}
  \item Set objectives: What are we fighting for?
  \item Allocate forces: Where does each unit go?
  \item Make decisions: When to attack, retreat, or commit reserves
  \item Inspire: Rally troops, maintain morale
  \item Adapt: Respond to changing conditions
\end{enumerate}

\subsection{9.2 Commander Actions}

\subsubsection{9.2.1 Command Roll}

As in Chapter 3, commanders make Command rolls (Presence + Tactics) to issue effective orders.

\textbf{Outcomes:}
\begin{itemize}
  \item Critical Success: All units gain +1 Position this round
  \item Success: Units act normally
  \item Partial: One unit acts at +1 Position, one at -1 Position
  \item Miss: All units act at -1 Position
\end{itemize}

\textbf{Spending Boons:} Commander may spend 1 Boon to:
\begin{itemize}
  \item Reroll Command roll
  \item Grant +1 Position to one unit this round
  \item Grant +2 Morale to one unit for one round
\end{itemize}

\subsubsection{9.2.2 Rally}

\textbf{Action:} Presence + Command\\
\textbf{Effect:} Restore Cohesion to a routing or shaken unit (see Chapter 8)

\subsubsection{9.2.3 Inspiring Speech}

\textbf{Action:} Presence + Performance\\
\textbf{Effect:} All units within hearing range gain +1 Morale for the duration of the speech (one round). A critical success grants +2 Morale.

\subsubsection{9.2.4 Tactical Assessment}

\textbf{Action:} Wits + Tactics\\
\textbf{Effect:} Gain insight into enemy dispositions. Ask the GM one question about the enemy (e.g., ``Where are their reserves?'' ``What is their commander's plan?'').

\subsubsection{9.2.5 Decisive Intervention}

\textbf{Action:} Physical (melee combat)\\
\textbf{Effect:} Commander enters combat personally. This may inspire troops (+1 Morale) but also risks the commander's life (if the commander is incapacitated, the army suffers -2 Morale).

\subsection{9.3 Heroes on the Battlefield}

PCs in mass combat are heroes. They can:
\begin{enumerate}
  \item Lead units: Command a unit as its leader
  \item Fight as champions: Engage enemy champions or dangerous units
  \item Perform special actions: Sabotage, scouting, magical support
  \item Change the battle: Do something that affects the course of the battle
\end{enumerate}

\subsubsection{9.3.1 Champion Actions}

\textbf{Duel:} Engage an enemy commander or champion.
\begin{itemize}
  \item Resolve as standard combat
  \item The outcome affects morale for both sides
\end{itemize}

\textbf{Sabotage:} Disrupt enemy operations.
\begin{itemize}
  \item Roll appropriate skill (Stealth, Craft, etc.)
  \item Effect: Enemy loses a Resource, suffers a Morale penalty, or loses a key unit
\end{itemize}

\textbf{Scout:} Gather intelligence.
\begin{itemize}
  \item Roll Wits + Survival (or appropriate)
  \item Effect: Gain +1 Position on the next Command roll
\end{itemize}

\textbf{Magic:} Use magical abilities to affect the battle.
\begin{itemize}
  \item Roll appropriate skill
  \item Effect: Area attacks, buffs, debuffs, communication, etc.
\end{itemize}

\subsubsection{9.3.2 Heroic Success}

When a hero succeeds in an action, they may affect the mass combat.

\textbf{Effects (choose one):}
\begin{itemize}
  \item Rally: Restore 2 Cohesion to a friendly unit
  \item Demoralize: Give an enemy unit -1 Morale
  \item Breakthrough: Open a gap in the enemy line
  \item Heroic Inspiration: All allies gain +1 Position for one round
\end{itemize}

\subsection{9.4 The Friction of War}

War is chaotic. Plans fail. Heroes fall. Opportunities are missed.

\textbf{Friction Table (d6)}

When the GM spends SB or a unit fails an action, roll on this table:

\begin{tabular}{|l|l|}
\hline
d6 & Friction Event \\
\hline
1 & Miscommunication: Orders are misunderstood. Unit acts differently than commanded. \\
\hline
2 & Delay: A unit is late to arrive. Reduce its Cohesion by 1 (waiting is exhausting). \\
\hline
3 & Accident: A unit suffers casualties from friendly fire or a collapsed bridge. Lose 1 Cohesion. \\
\hline
4 & Rumor: A rumor of defeat spreads. All units suffer -1 Morale until the commander rallies them. \\
\hline
5 & Equipment Failure: A unit loses its advantage (archers run out of arrows, a ladder breaks). \\
\hline
6 & Commander Down: The unit's commander is incapacitated. Unit acts at -1 Position until a new commander is appointed. \\
\hline
\end{tabular}

\section{Chapter 10: Siege Warfare}

\subsection{10.1 The Siege Framework}

A siege is a sustained operation to take a fortified position. It involves:
\begin{itemize}
  \item Investment: Surrounding the fortification
  \item Bombardment: Destroying walls and defenses
  \item Assault: Attacking the walls
  \item Breach: Breaking through
  \item Occupation: Taking the fortification
\end{itemize}

\subsection{10.2 Siege Timers}

Sieges are tracked by timers that represent progress.

\textbf{Siege Timer (10):} Tracks overall siege progress.
\begin{itemize}
  \item Advances when the besiegers make a significant action (bombardment, assault, etc.)
  \item Retrets when the defenders make a successful counter-action
  \item When it fills, the siege is resolved (attacker wins)
\end{itemize}

\textbf{Defender Timer (10):} Tracks the defenders' resilience.
\begin{itemize}
  \item Advances when the defenders take casualties, morale drops, or supply runs low
  \item Retreats when the defenders rally or receive relief
  \item When it fills, the defenders surrender
\end{itemize}

\textbf{Supply Timer (8):} Tracks food and water.
\begin{itemize}
  \item Advances each day of siege
  \item Advances faster when the defenders are numerous or the supply is limited
  \item When it fills, the defenders begin to starve
\end{itemize}

\textbf{Breach Timer (6):} Tracks progress toward creating a breach in the fortifications.
\begin{itemize}
  \item Advances when mining or bombardment weakens the walls
  \item Retreats when defenders repair damage
  \item When it fills, a breach is made and assault actions gain +2 Effect
\end{itemize}

\subsection{10.3 Siege Actions}

\subsubsection{10.3.1 Attacker Actions}

\textbf{Bombard:} Use siege engines to damage walls and defenders.
\begin{itemize}
  \item Roll Artillery + Resources
  \item Success: Damage walls; defender casualties; advance Siege Timer
  \item Partial: Damage walls; limited casualties
  \item Miss: Damage limited; Resources lost; no advance
\end{itemize}

\textbf{Assault:} Attack the walls directly.
\begin{itemize}
  \item Roll Melee + Cap
  \item Success: Gain foothold; advance Siege Timer by 2
  \item Partial: Gain foothold; lose a unit
  \item Miss: Repulsed; lose a unit
\end{itemize}

\textbf{Mine:} Tunnel under walls.
\begin{itemize}
  \item Roll Craft + Tactics
  \item Success: Create breach; advance Siege Timer
  \item Partial: Tunnel collapses; lose a Resource
  \item Miss: Tunnel discovered; counter-mined
\end{itemize}

\textbf{Negotiate:} Offer terms to the defender.
\begin{itemize}
  \item Roll Presence + Sway
  \item Success: Defenders consider terms; lower Defender Timer
  \item Partial: Terms considered but not accepted
  \item Miss: Terms insulted; defender morale increases
\end{itemize}

\textbf{Relieve:} Break a siege.
\begin{itemize}
  \item Roll Melee + Tactics
  \item Success: Relief force arrives; siege is broken
  \partial: Relief force is delayed; both sides lose a unit
  \item Miss: Relief force is defeated; siege continues
\end{itemize}

\subsubsection{10.3.2 Defender Actions}

\textbf{Repair:} Fix damaged walls.
\begin{itemize}
  \item Roll Craft + Resources
  \item Success: Walls repaired; Siege Timer retreats
  \item Partial: Partial repair; limited effect
  \item Miss: No repair; resources lost
\end{itemize}

\textbf{Sortie:} Attack the besiegers.
\begin{itemize}
  \item Roll Melee + Cap
  \item Success: Cause casualties; slow siege progress
  \item Partial: Cause casualties; lose a unit
  \item Miss: Repulsed; lose a unit
\end{itemize}

\textbf{Counter-Mine:} Detect and collapse enemy mines.
\begin{itemize}
  \item Roll Wits + Tactics
  \item Success: Enemy mine destroyed
  \item Partial: Mine detected but not destroyed
  \item Miss: Mine collapses your own walls
\end{itemize}

\textbf{Rally:} Restore defender morale.
\begin{itemize}
  \item Roll Presence + Command
  \item Success: Defender morale restored; Defender Timer retreats
  \item Partial: Partial restoration
  \item Miss: Morale falls further
\end{itemize}

\subsection{10.4 Starvation and Surrender}

When the Supply Timer fills, the defenders begin to starve.

\textbf{Starvation Effects:}
\begin{itemize}
  \item -1 Cohesion per day
  \item -1 to all rolls (physical, mental, and magical)
  \item Higher chance of surrender
\end{itemize}

\textbf{Surrender Conditions:}
\begin{itemize}
  \item Supply Timer at 0
  \item Siege Timer at 10
  \item Defender Timer at 0
  \item Commander killed
\end{itemize}

\textbf{Surrender Outcomes:}
\begin{itemize}
  \item Negotiated: Terms are offered; defenders may leave with honor or be absorbed into the attacker's forces
  \item Unconditional: Defenders surrender without terms; may be imprisoned, enslaved, or killed
\end{itemize}

\section{Part IV: The War Beyond the Battle}

\section{Chapter 11: Logistics—The Sinews of War}

\subsection{11.1 The Logistics System}

Logistics is the art of moving, feeding, and supplying an army. It is the least glamorous aspect of war and arguably the most important.

\textbf{Logistics Tracks:}
\begin{itemize}
  \item Food: Rations, water, animal feed
  \item Arms: Weapons, ammunition, armor
  \item Materials: Tents, tools, siege equipment, magical supplies
\end{itemize}

\textbf{Each Track is a Timer:}
\begin{itemize}
  \item Food [10]: Advances each day; retreats when supplies are replenished
  \item Arms [10]: Advances when equipment is lost or used; retreats when resupplied
  \item Materials [10]: Advances when siege equipment or magical supplies are used; retreats when replaced
\end{itemize}

\subsection{11.2 Supply Sources}

Armies acquire supplies from:

\textbf{Local Foraging:} Living off the land.
\begin{itemize}
  \item Roll Survival + Cap
  \item Success: Gain 1 Food
  \item Partial: Gain 1 Food but lose 1 Cohesion (annoying the locals)
  \item Miss: Gain nothing; may provoke local opposition
\end{itemize}

\textbf{Supply Lines:} Bringing supplies from friendly territory.
\begin{itemize}
  \item Requires secure routes; vulnerable to raiding
  \item Roll Command + Logistics (if commander is handling supply)
  \item Success: Supplies arrive; all tracks retreat
  \item Partial: Supplies arrive but damaged; retreat one track
  \item Miss: Supplies delayed; tracks advance
\end{itemize}

\textbf{Aquisition:} Stealing from the enemy.
\begin{itemize}
  \item Requires raiding parties
  \item Roll Melee + Cap (for raiding party)
  \item Success: Gain 2 Food (or Arms/Materials)
  \item Partial: Gain 1 Food, lose a unit
  \item Miss: Gain nothing; lose a unit
\end{itemize}

\subsection{11.3 Logistics Effects}

\begin{tabular}{|l|l|}
\hline
Track Level & Effect \\
\hline
0–2 & Fully supplied; no penalty \\
\hline
3–5 & Short rations; -1 Cohesion per day \\
\hline
6–8 & Low supplies; -2 Cohesion per day; -1 to all rolls (physical) \\
\hline
9–10 & Emergency! Marching, fighting, or digging may require a Cohesion check \\
\hline
\end{tabular}

\subsection{11.4 Logistics and Morale}

The connection between supplies and morale is direct. Soldiers who are hungry, cold, or poorly equipped will not fight well.

\textbf{Mechanical Effect:} For every 2 levels of a Logistics track that are above 3, the army suffers -1 Morale.

\section{Chapter 12: Morale, Cohesion, and the Human Element}

\subsection{12.1 Morale Factors}

Morale is affected by:

\begin{tabular}{|l|l|}
\hline
Factor & Effect \\
\hline
Victory & +1 (per significant victory) \\
\hline
Defeat & -1 (per significant defeat) \\
\hline
Supply & +1 (well-supplied) to -2 (starving) \\
\hline
Command & +1 (good) to -2 (bad) \\
\hline
Pay & +1 (paid) to -2 (unpaid) \\
\hline
Cause & +1 (believed in) to -1 (disbelieved) \\
\hline
Elite Unit & +1 (permanent) \\
\hline
Green Unit & -1 (permanent) \\
\hline
\end{tabular}

\textbf{Morale Check:} A unit with Morale 3 or lower must make a Morale check when ordered to advance or when facing a significant threat. Failure causes the unit to hesitate or retreat.

\subsection{12.2 Cohesion in Combat}

Cohesion represents the unit's discipline and unity.

\textbf{Cohesion Effects:}
\begin{itemize}
  \item 5–6: Battle-ready; no penalty
  \item 3–4: Shaken; -1 to all rolls
  \item 1–2: Brittle; -2 to all rolls; must make Morale check when attacked
  \item 0: Broken; unit routs
\end{itemize}

\textbf{Restoring Cohesion:}
\begin{itemize}
  \item Rally (Command Action): Restores 1–2 Cohesion
  \item Rest: Restores 1 Cohesion per day
  \item Victory: Restores 1 Cohesion
  \item Supplies: Restores 1 Cohesion if supply situation improves
\end{itemize}

\subsection{12.3 The Human Element—Stories of Soldiers}

\textit{``I have seen a man fight through a wound that would have killed a lesser man. I have seen a man break when a leaf brushed his shoulder. There is no logic in the human heart. Only stories.''}

\textbf{Soldier Stories (d10):} When a unit is in combat, roll on this table to discover something about the soldiers in it.

\begin{tabular}{|l|l|}
\hline
d10 & Story \\
\hline
1 & The Reluctant Soldier: A soldier who joined to protect their family. They fight with desperate courage. \\
\hline
2 & The Veteran: A soldier who has seen too much. They are cynical, unbreakable, and loyal to their friends. \\
\hline
3 & The Dreamer: A soldier who believes in a cause. They fight with idealism. \\
\hline
4 & The Refugee: A soldier who lost everything to war. They fight for a new home. \\
\hline
5 & The Conscript: A soldier who was forced into service. They fight because they have to. \\
\hline
6 & The Mercenary: A soldier who fights for coin. They are professional, but not loyal. \\
\hline
7 & The Zealot: A soldier who believes their cause is divine. They fight with fanaticism. \\
\hline
8 & The Coward: A soldier who is afraid. They may flee or they may surprise everyone with courage. \\
\hline
9 & The Hero: A soldier who has the potential to be a legend. They are the one the songs will be sung about. \\
\hline
10 & The Ghost: A soldier who has already died in spirit. They fight because they have nothing left to lose. \\
\hline
\end{tabular}

\section{Chapter 13: Campaign Warfare—War as a Series of Battles}

\subsection{13.1 Campaign Structure}

A campaign is a series of battles, maneuvers, and strategic decisions. It has a narrative arc.

\textbf{Campaign Phases:}
\begin{enumerate}
  \item Opening Moves: Strategy, logistics, and initial engagements
  \item Maneuver: Flanking, feints, and jockeying for position
  \item Climax: Major battle that decides the campaign
  \item Resolution: Pursuit, negotiations, or continuation
\end{enumerate}

\subsection{13.2 Campaign Turns}

Campaign turns represent periods of time (days, weeks, months) during which strategic decisions are made.

\textbf{Turn Sequence:}
\begin{enumerate}
  \item Planning Phase: Decide strategic objectives and movement of forces
  \item Movement Phase: Move armies; resolve marches
  \item Combat Phase: Resolve battles and sieges
  \item Logistics Phase: Manage supplies, recruitment, and replacements
  \item End Phase: Update status; determine new objectives
\end{enumerate}

\subsection{13.3 Campaign Decisions}

The PCs make campaign-level decisions that affect the war.

\textbf{Decision Types:}
\begin{itemize}
  \item Strategic: ``Where do we march?'' ``Who do we ally with?''
  \item Tactical: ``How do we use our forces?'' ``Where do we strike?''
  \item Logistical: ``Where do we get supplies?'' ``How do we move them?''
  \item Political: ``What do we promise our allies?'' ``What do we demand of our enemies?'
\end{itemize}

\textbf{Decision Effects:}
\begin{itemize}
  \item Success: Gain advantage in the next battle; alliances hold; supplies flow
  \item Partial: Mixed results; both sides gain advantage; supplies are strained
  \item Failure: Disadvantage in the next battle; alliances fray; supplies are cut
\end{itemize}

\subsection{13.4 Campaign Timers}

Campaign timers represent larger strategic pressures.

\begin{tabular}{|l|l|}
\hline
Timer & Tracks \\
\hline
War Weariness [10] & Exhaustion of the war effort; when full, both sides seek peace or collapse \\
\hline
Alliance Stability [8] & Cohesion of the PC's alliances; when full, alliances break \\
\hline
Enemy Momentum [8] & Enemy's strategic advantage; when full, the enemy may achieve a significant victory \\
\hline
Supply Chain [10] & Overall supply situation; when full, the army is starving and near collapse \\
\hline
\end{tabular}

\textbf{Advancing Timers:}
\begin{itemize}
  \item War Weariness: Advances with each major battle; retreats when victory is achieved
  \item Alliance Stability: Advances when promises are broken or key battles are lost; retrets when alliances are reinforced
  \item Enemy Momentum: Advances when the enemy gains a victory or the PCs fail a strategy; retreats when the PCs achieve a victory
  \item Supply Chain: Advances when supply lines are cut or resources are scarce; retreats when supplies are secured
\end{itemize}

\subsection{13.5 The Art of the Campaign}

\textit{``A war is not won by a single battle. It is won by a thousand decisions—each one small, each one hard, each one made by a person who does not know if they are making the right choice.''}

\section{Chapter 14: Aftermath—Victory, Defeat, and the Scars of War}

\subsection{14.1 The Cost of Victory}

Victory is not free. Even winning a war leaves scars.

\textbf{Victory Costs:}
\begin{itemize}
  \item Casualties: The dead, the wounded, the traumatized
  \item Resources: Money, food, equipment
  \item Morale: The army is tired, the country is tired, the allies are tired
  \item Political: New enemies are made; old friends may be lost
  \item Scars: The land is scarred; the people are scarred; the soldiers are scarred
\end{itemize}

\textbf{Mechanical Effects:}
\begin{itemize}
  \item Casualties: Units lose Cohesion; replacements take time
  \item Resources: Supplies must be replenished; money must be found
  \item Morale: The army and the country need time to recover
  \item Political: Relationships with allies and enemies may change
  \item Scars: The land, the people, and the soldiers are changed
\end{itemize}

\subsection{14.2 Defeat}

Defeat is devastating. It means losing people, losing resources, and losing hope.

\textbf{Defeat Costs:}
\begin{itemize}
  \item Casualties: More severe than in victory
  \item Resources: Lost or captured
  \item Morale: Severely damaged
  \item Political: Allies may abandon you; enemies may press their advantage
  \item Scars: Defeat leaves deep scars that may never heal
\end{itemize}

\textbf{Defeat Effects:}
\begin{itemize}
  \item Units: Lost or shattered
  \item Resources: Lost or captured
  \item Morale: Severely damaged
  \item Political: Allies may abandon you; enemies may press their advantage
\end{itemize}

\subsection{14.3 The Scars of War}

War leaves scars—on the land, on the people, and on the soldiers.

\textbf{Land Scars:}
\begin{itemize}
  \item Ruin: Cities, towns, and villages are destroyed
  \item Famine: Farms are burned; crops are ruined
  \item Waste: Land is poisoned by war; the earth is scarred by trenches and tunnels
  \item Magic: Magical use can scar the land permanently
\end{itemize}

\textbf{People Scars:}
\begin{itemize}
  \item Trauma: Soldiers and civilians are traumatized by war
  \item Grief: Loss of family, friends, and homes
  \item Displacement: Refugees are displaced
  \item Anger: Anger and hatred grow
\end{itemize}

\textbf{Soldier Scars:}
\begin{itemize}
  \item Injury: Physical wounds that may not fully heal
  \item Trauma: Psychological wounds that may never heal
  \item Change: Soldiers are changed by war—often permanently
  \item Disconnection: Soldiers may not fit back into civilian life
\end{itemize}

\subsection{14.4 Returning Home}

\textbf{Mechanics for Returning Soldiers:}
\begin{itemize}
  \item Rolling Scars: After a campaign, each soldier makes a Spirit + Resolve roll. Success: the soldier is unchanged. Partial: the soldier gains a permanent Condition. Miss: the soldier gains a serious Condition or must leave the campaign.
\end{itemize}

\textbf{Player Characters:}
\begin{itemize}
  \item Gain a Condition: After a campaign, PCs may gain a Condition (e.g., ``War-Weary,'' ``Haunted,'' ``Hardened'').
  \item Gain a Boon: They may also gain a Boon (e.g., ``Combat Veteran,'' ``Network of Contacts,'' ``Heroic Reputation'').
\end{itemize}

\textbf{Scar Mechanics (Optional):}
\begin{tabular}{|l|l|}
\hline
Scar & Mechanical Effect \\
\hline
War-Weary & -1 on Spirit rolls to resist fear; +1 on Endurance rolls (used to hardship) \\
\hline
Haunted & Nightmares: -1 on Wits rolls after a failed rest; +1 on Insight rolls (hypervigilance) \\
\hline
Hardened & -1 on social rolls requiring empathy; +1 on Command rolls \\
\hline
Bonded Survivors & +1 when fighting alongside a specific unit or character; -1 when they are threatened \\
\hline
\end{tabular}

\section{Appendices}

\section{Appendix A: Quick Reference Tables}

\subsection{A.1 Command Roll Outcomes}

\begin{tabular}{|l|l|}
\hline
Outcome & Effect \\
\hline
Critical Success & All units gain +1 Position; commander gains 1 Boon \\
\hline
Success & Units act normally; one unit may act at +1 Position \\
\hline
Partial & One unit at +1 Position, one at -1 Position \\
\hline
Miss & All units at -1 Position; commander may spend 1 Boon to negate \\
\hline
\end{tabular}

\subsection{A.2 Morale Check}

\textbf{Roll:} Cap + Command + Morale Bonuses - Penalties

\textbf{Outcomes:}
\begin{itemize}
  \item Success: Unit holds
  \item Partial: Unit is shaken (-1 Position next turn)
  \item Miss: Unit routs
\end{itemize}

\subsection{A.3 Unit Harm}

\begin{tabular}{|l|l|}
\hline
Cap & Harm Levels \\
\hline
1 & 2 \\
\hline
2 & 3 \\
\hline
3 & 4 \\
\hline
4 & 5 \\
\hline
5 & 6 \\
\hline
\end{tabular}

\subsection{A.4 Morale Modifiers}

\begin{tabular}{|l|l|}
\hline
Factor & Effect \\
\hline
Victory & +1 \\
\hline
Defeat & -1 \\
\hline
Supply & +1 to -2 \\
\hline
Command & +1 to -2 \\
\hline
Pay & +1 to -2 \\
\hline
Cause & +1 to -1 \\
\hline
Elite Unit & +1 \\
\hline
Green Unit & -1 \\
\hline
\end{tabular}

\subsection{A.5 Logistics Tracks}

\begin{tabular}{|l|l|}
\hline
Track & Effect at 6+ \\
\hline
Food & -1 Cohesion/day \\
\hline
Arms & -1 to all attack rolls \\
\hline
Materials & -1 to all siege/magic rolls \\
\hline
\end{tabular}

\subsection{A.6 Terrain Quick Reference}

\begin{tabular}{|l|l|l|l|l|}
\hline
Terrain & Movement & Cover & Vision & Combat \\
\hline
Open Field & Normal & None & Good & Normal \\
\hline
Hill & 1.5x up & Partial & Varies & +1 Position \\
\hline
Forest & 0.5x & Good & Poor & Ambush \\
\hline
Marsh & 0.25x & None & Poor & -1 Position \\
\hline
Urban & 0.5x & Good & Limited & +1 Defense \\
\hline
Water & 0.5x & None & Poor & Vulnerable \\
\hline
\end{tabular}

\section{Appendix B: Scenario Generator}

\subsection{B.1 Battle Generator (d12)}

\begin{tabular}{|l|l|l|}
\hline
d12 & Battle Type & Description \\
\hline
1 & Field Battle & Two armies meet on open ground \\
\hline
2 & Raid & A quick strike against a vulnerable target \\
\hline
3 & Ambush & An attack from concealment \\
\hline
4 & Siege & Attack or defense of a fortified position \\
\hline
5 & River Crossing & An attempt to cross a river under fire \\
\hline
6 & Night Battle & A battle fought in darkness \\
\hline
7 & Fog Battle & A battle fought in fog or mist \\
\hline
8 & Rearguard Action & A delaying action to allow retreat \\
\hline
9 & Pursuit & A chase after a defeated enemy \\
\hline
10 & Breakthrough & An attempt to break through enemy lines \\
\hline
11 & Relief & An attempt to break a siege \\
\hline
12 & Naval Battle & A battle fought on water \\
\hline
\end{tabular}

\subsection{B.2 Battle Objective Generator (d10)}

\begin{tabular}{|l|l|}
\hline
d10 & Objective \\
\hline
1 & Destroy the Enemy & Annihilate the opposing force \\
\hline
2 & Hold the Line & Stop the enemy advance \\
\hline
3 & Capture a Key Objective & Take a hill, bridge, or fortification \\
\hline
4 & Protect a Person & Rescue or defend a VIP \\
\hline
5 & Gain Intelligence & Capture prisoners or documents \\
\hline
6 & Buy Time & Delay the enemy to allow others to escape \\
\hline
7 & Break Through & Open a route to friendly forces \\
\hline
8 & Destroy Supplies & Raid enemy supply lines \\
\hline
9 & Show of Force & Demonstrate strength to intimidate \\
\hline
10 & Revenge & Settle a score; prove a point \\
\hline
\end{tabular}

\subsection{B.3 Battle Twist Generator (d10)}

\begin{tabular}{|l|l|}
\hline
d10 & Twist \\
\hline
1 & Reinforcements & One side receives unexpected reinforcements \\
\hline
2 & Weather & Weather changes dramatically \\
\hline
3 & Commander Down & A key commander is incapacitated or killed \\
\hline
4 & Betrayal & A unit switches sides \\
\hline
5 & Supply Issue & A supply line is cut or destroyed \\
\hline
6 & Heroism & A hero emerges to change the course of battle \\
\hline
7 & Panic & A unit routs unexpectedly \\
\hline
8 & Terrain & Terrain changes (e.g., a bridge collapses, a fire starts) \\
\hline
9 & Message & A message arrives that changes the strategic situation \\
\hline
10 & Omen & A divine or magical sign appears \\
\hline
\end{tabular}

\section{Appendix C: Unit Templates}

\subsection{C.1 Infantry Units}

\subsubsection{Levy Infantry}

\textbf{Unit Name:} [Region] Levy\\
\textbf{Type:} Light Infantry\\
\textbf{Cap:} 1\\
\textbf{Scale:} II (Company, 50–100 men)\\
\textbf{Speed:} Standard\\
\textbf{Tags:} Green\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6]

\subsubsection{Standard Infantry}

\textbf{Unit Name:} [Region] Infantry\\
\textbf{Type:} Medium Infantry\\
\textbf{Cap:} 2\\
\textbf{Scale:} II (Company, 50–100 men)\\
\textbf{Speed:} Standard\\
\textbf{Tags:} None\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\subsubsection{Heavy Infantry}

\textbf{Unit Name:} [Region] Heavy Infantry\\
\textbf{Type:} Heavy Infantry\\
\textbf{Cap:} 3\\
\textbf{Scale:} II (Company, 50–100 men)\\
\textbf{Speed:} Slow\\
\textbf{Tags:} Shields, Armor\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\subsection{C.2 Cavalry Units}

\subsubsection{Light Cavalry}

\textbf{Unit Name:} [Region] Light Cavalry\\
\textbf{Type:} Light Cavalry\\
\textbf{Cap:} 2\\
\textbf{Scale:} II (Squadron, 30–50)\\
\textbf{Speed:} Fast\\
\textbf{Tags:} Skirmish\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\subsubsection{Heavy Cavalry}

\textbf{Unit Name:} [Region] Heavy Cavalry\\
\textbf{Type:} Heavy Cavalry\\
\textbf{Cap:} 3\\
\textbf{Scale:} II (Squadron, 30–50)\\
\textbf{Speed:} Fast\\
\textbf{Tags:} Armor, Charge\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\subsection{C.3 Skirmisher Units}

\subsubsection{Archers}

\textbf{Unit Name:} [Region] Archers\\
\textbf{Type:} Archers\\
\textbf{Cap:} 2\\
\textbf{Scale:} II (Company, 50–100)\\
\textbf{Speed:} Standard\\
\textbf{Tags:} Ranged\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\subsubsection{Crossbowmen}

\textbf{Unit Name:} [Region] Crossbowmen\\
\textbf{Type:} Crossbowmen\\
\textbf{Cap:} 3\\
\textbf{Scale:} II (Company, 50–100)\\
\textbf{Speed:} Slow\\
\textbf{Tags:} Ranged, Armor-Piercing\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\subsection{C.4 Special Units}

\subsubsection{Siege Engineers}

\textbf{Unit Name:} [Region] Engineers\\
\textbf{Type:} Engineers\\
\textbf{Cap:} 2\\
\textbf{Scale:} I (Squad, 10–20)\\
\textbf{Speed:} Slow\\
\textbf{Tags:} Siege\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\subsubsection{Artillery}

\textbf{Unit Name:} [Region] Artillery\\
\textbf{Type:} Artillery\\
\textbf{Cap:} 2\\
\textbf{Scale:} I (Squad, 10–20)\\
\textbf{Speed:} Very Slow\\
\textbf{Tags:} Siege, Area\\
\textbf{Tracks:} Cohesion [6], Exposure [6], Resources [6

\section{Appendix D: War Campaign Generator}

\subsection{D.1 Campaign Structure Generator (d10)}

\begin{tabular}{|l|l|l|}
\hline
d10 & Campaign Type & Description \\
\hline
1 & Border War & Small-scale conflict over a border dispute \\
\hline
2 & Civil War & Internal conflict between factions \\
\hline
3 & Invasion & One nation invades another \\
\hline
4 & Defensive War & A nation fights to defend itself \\
\hline
5 & Holy War & A conflict driven by religious fervor \\
\hline
6 & War of Succession & Conflict over who rules \\
\hline
7 & War of Independence & A region fights to break away \\
\hline
8 & Proxy War & Two larger powers fight by proxy \\
\hline
9 & Total War & An existential conflict \\
\hline
10 & War of Conquest & One nation aims to conquer another \\
\hline
\end{tabular}

\subsection{D.2 Campaign Phase Generator (d6)}

\begin{tabular}{|l|l|l|}
\hline
d6 & Phase & What Happens \\
\hline
1 & Opening Moves & Initial engagements; both sides test each other \\
\hline
2 & Maneuver & Armies move to gain advantage \\
\hline
3 & Climax & A decisive battle is fought \\
\hline
4 & Stalemate & Neither side can gain advantage \\
\hline
5 & Pursuit & One side pursues a defeated enemy \\
\hline
6 & Negotiation & Both sides seek a diplomatic solution \\
\hline
\end{tabular}

\subsection{D.3 Campaign Twist Generator (d10)}

\begin{tabular}{|l|l|}
\hline
d10 & Twist \\
\hline
1 & Third Party & A new faction enters the war \\
\hline
2 & Betrayal & An ally betrays the PCs \\
\hline
3 & Unexpected Alliance & Former enemies become allies \\
\hline
4 & Plague & An outbreak affects the armies \\
\hline
5 & Miracle & A divine or magical event turns the tide \\
\hline
6 & Supply Crisis & A major supply line is cut \\
\hline
7 & Mutiny & A unit rebels against its commanders \\
\hline
8 & Dynastic Crisis & A succession crisis disrupts one side \\
\hline
9 & Peace Offer & The enemy offers terms \\
\hline
10 & Daring Raid & A raid achieves an unexpected outcome \\
\hline
\end{tabular}

\newpage
\section{Campaign Seeds for \textit{The Art of War}}

The following campaign seeds are original constructs built from the strategic geometry of \textit{Fate's Edge}. They are intended for use with the mass combat, logistics, and command systems presented in \textit{The Art of War} expansion. All proper nouns have been abstracted; only the underlying patterns of war remain.

Each seed is presented as a multi-arc campaign with clearly defined faction timers, pivotal strategic decisions, and multiple possible conclusions. The Game Master should map the geographic elements (rivers, mountain passes, fortified cities, etc.) onto the existing map of the Amaranthine world, the Violet Steppe, or the contested borderlands of the fractured empire.

\subsection{The Siege of the White City}

\textbf{Premise:} For three years, the Everflame's Crusade has swept across the southern deserts, burning temples and breaking the old covenants. Now the holy armies of the Flame—a coalition of Ecktorian knights, fanatical pilgrims, and Black Banner mercenaries—have surrounded the White City, the last bastion of the desert lords. The city's wells are failing. Its walls are cracked. Its defenders are outnumbered ten to one.

But the commander of the Crusade is no fanatic. He is a veteran of the Ashaan frontier, a man who has seen what zeal does to men and cities. And the defender of the White City is no desperate fanatic either—he is a prince of the old blood, a poet and a warrior, who knows that victory is impossible but surrender is worse.

The siege is a war of brinkmanship: raids, counter-raids, parleys, and the slow, grinding calculus of starvation. The Crusaders cannot storm the walls without catastrophic losses. The defenders cannot hold out through another winter. Both sides are waiting for the other to break—or for a miracle.

\subsubsection{Factions \& Actors}

\begin{itemize}
    \item \textbf{The Crusader Host:} The Everflame's army—a polyglot force of Ecktorian heavy infantry, fanatical pilgrim levies, and Black Banner mercenaries bound by pay and plunder. Their siege engines are the finest in the known world. \textit{Strategic Goal:} Take the city, break the last organized resistance, and open the southern trade routes to Everflame dominion. \textit{Internal Tension:} The mercenaries are restless without pay; the pilgrims demand blood; the veterans counsel patience.
    \item \textbf{The White City Garrison:} The defenders—a mixed force of Fharan desert cavalry, Ashaani Immortal veterans (exiles who fled the Veiled Sisters' coup), and the city's own militia. They are outnumbered, undersupplied, and running out of hope. But they know the desert, and they know their walls. \textit{Strategic Goal:} Hold out until the winter rains make the Crusaders' supply lines untenable, or until a relief force from the eastern satrapies arrives. \textit{Internal Tension:} Some in the garrison want to surrender on honorable terms; others would rather die fighting.
    \item \textbf{The Neutral Merchants:} A guild of spice traders and financiers, operating from the city's Grand Bazaar, who have maintained cordial relations with both sides throughout the siege. They want the fighting to stop—so trade can resume. They are the only ones with a foot in both camps. \textit{Strategic Goal:} Broker a truce, preserve their warehouses, and position themselves as the indispensable intermediaries in any peace.
    \item \textbf{The Malachai Covenant (Secret Faction):} A shadowy sect within the city that worships Malachai, the Chained Angel. They believe the siege is a divine test—and that the only way to save the city is to sacrifice its prince to the Well of Ichor. Among their number is Baron Quentin Asselin himself, a man who once drank the angel's blood and has been seeking a way to undo that choice ever since. \textit{Strategic Goal:} Ensure the prince is delivered to Malachai—willingly or not—and thus break the siege through supernatural means.
\end{itemize}

\subsubsection{Strategic Timers}

\begin{itemize}
    \item \textbf{Siege Intensity [8]:} Measures the pressure of the Crusaders' operations. Increases after each successful assault, breach, or capture of an outer wall; decreases when the Crusaders suffer a reverse, a supply convoy is cut, or a parley extends a truce. At 4, the outer walls are compromised (defenders suffer -1 Position on all defensive rolls); at 8, the inner city is breached.
    \item \textbf{Defender Morale [6]:} Reflects the garrison's will to continue. Increases when a raid is repelled, a relief force is spotted, or a victory is won; decreases when food runs short, a key officer is killed, or a spy is uncovered. At 0, the garrison may mutiny or surrender; at 6, it fights with desperate valor (+1 Position to all defensive actions).
    \item \textbf{Relief Column [6]:} The approach of an allied army from the eastern satrapies, sent to break the siege. Increases with each day the defenders hold out; decreases if the Crusaders intercept and destroy the column. At 6, the relief force arrives—and the siege becomes a battle of two armies.
\end{itemize}

\subsubsection{Campaign Arcs}

\begin{enumerate}
    \item \textbf{The Investment:} The Crusaders complete the encirclement of the White City, cutting off its water and supply lines. The defenders mount desperate sorties to seize water from the besiegers' wells. \textit{Key Decision:}
    \begin{itemize}
        \item \textit{Fortify the Inner City:} Pull the garrison back to the inner walls, sacrificing the outer districts to conserve manpower and supplies (lose 2 segments of Siege Intensity but gain +1 Morale for the defenders; the outer city is sacked).
        \item \textit{Hold the Outer Walls:} Commit the garrison to a desperate defense of the entire perimeter, hoping to bleed the Crusaders dry (increase Siege Intensity by 2 this round, but if successful, Crusader Morale takes a severe blow).
    \end{itemize}
    \item \textbf{The Parley:} A lull in the fighting. The Crusader commander and the defending prince meet under a flag of truce outside the walls. They do not negotiate peace—they negotiate the terms of the city's surrender should it fall. \textit{Key Decision:}
    \begin{itemize}
        \item \textit{Generous Terms:} The prince offers to surrender in exchange for the safe passage of all citizens, the preservation of the city's shrines, and a generous tribute (requires a successful Presence+Sway roll against a DV set by the Crusader commander's own internal pressures; success secures a temporary ceasefire and buys time for relief; failure hardens the Crusaders' resolve).
        \item \textit{Defiant Stance:} The prince refuses to even discuss surrender, demanding the Crusaders withdraw or face a fight to the last stone (increases Defender Morale by 2 but also increases Siege Intensity by 1; the Crusader commander respects the prince's courage but is now committed to total victory).
    \end{itemize}
    \item \textbf{The Breach:} A section of the inner wall collapses under concentrated bombardment. The Crusaders pour through the gap. The defenders mount a last stand in the streets. \textit{Key Decision:}
    \begin{itemize}
        \item \textit{Fight in the Streets:} The garrison mounts a desperate, house-to-house defense, hoping to break the Crusaders' momentum before the inner citadel falls (requires a Mass Combat roll with a -2 Position penalty; success forces the Crusaders back, buying time for relief; failure sees the city fall).
        \item \textit{Accept the Inevitable:} The prince orders the garrison to lay down their arms—not in surrender, but to spare the city from sack. He walks out to meet the Crusader commander, his sword held out, hilt first (ends the siege immediately; the city is spared mass slaughter, but the prince must endure the shame of capitulation and the Crusaders' demands).
    \end{itemize}
    \item \textbf{The Intervention (Optional Twist):} Baron Quentin Asselin, acting on behalf of the Malachai Covenant, attempts to kidnap the prince and deliver him to the Well of Ichor, believing that only a willing sacrifice can save the city. The party must protect the prince—or decide whether Malachai's "salvation" is worth the price.
\end{enumerate}

\subsubsection{Potential Endings}

\begin{itemize}
    \item \textbf{The Honorable Surrender:} The prince surrenders on generous terms. The city is spared sack, its people allowed to leave with their lives and a portion of their goods. The Crusader commander is hailed for his mercy—but the Everflame's zealots see it as a failure of faith. The prince goes into exile, becoming a wandering poet, forever lamenting the city he could not hold.
    \item \textbf{The Last Stand:} The city falls, but the fighting is so fierce that the Crusader army is gutted. The prince dies on the walls, his name becoming a rallying cry for the resistance. The Everflame's victory is pyrrhic; the southern trade routes remain closed for another generation.
    \item \textbf{The Malachai Gambit:} The prince is delivered to the Well of Ichor. Malachai's power manifests, shattering the Crusader lines in a night of terror. The siege is broken—but the city is now a festering wound, its prince cursed, its people under the shadow of the Chained Angel. A new evil has replaced the old.
    \item \textbf{The Stalemate:} Neither side can break the other. The Crusaders withdraw after winter sets in, promising to return. The city survives, but its walls are shattered, its wells dry. The prince and the Crusader commander meet once more, this time over a cup of tea in the city's ruins, and they speak not of war, but of the cost of war. A fragile peace, born of exhaustion, holds for a decade—until the cycle begins again.
\end{itemize}

\end{itemize}\subsection{The King of the Mountain}

\textbf{Premise:} An ancient Aeler king, the last of his line, has gathered a force of dwarven-stone warriors, human mercenaries, and loyal lowland allies to hold a vital mountain pass against a tide of eastern confederates pouring from the high steppe. The campaign is a brutal defensive struggle, a war of attrition where every yard of stone is bitterly contested, and the outcome hinges on engineering, terrain, and the unyielding will of the defenders—but the invaders are no mere rabble. They are a sophisticated, oath-bound coalition with their own strategic genius, their own grievances, and their own vision of what the world should become.

\subsubsection{Factions & Actors}
\begin{itemize}
\item \textbf{The Mountain Kingdom:} The Aeler defenders, reinforced by hired dwarven mercenaries (masters of fortification and siegecraft) and allied human levies from the lowlands (providing numbers and fresh blood). \textit{Strategic Goal:} Hold the pass until the winter snows force the invaders to turn back, inflicting maximal cost on the enemy while preserving the kingdom's independence.
\item \textbf{The Great Khanate:} A vast, nomadic confederation from the far eastern steppe, led by a warlord whose reputation for strategic genius is matched only by his mastery of steppe logistics and rapid maneuver. The Khanate is not a mindless horde; it is a coalition of clan-lords, each with their own interests, bound by oath-rings and the promise of plunder. Their warriors are superbly mounted, their intelligence network reaches deep into the lowlands, and their siege engineers—recruited from conquered hill-tribes—are the equal of any Aeler. \textit{Strategic Goal:} Break through the pass and secure winter grazing lands for the clans, who have been pushed eastward by a generation of drought and overpopulation. They do not seek to destroy—they seek to survive and, if possible, to settle.
\item \textbf{The Neutral Duchies:} Human lords who hold the fertile valleys and fortified towns just beyond the pass's mouth; they are terrified of the Khanate's approach but too cynical to throw their lot in with the Aeler cause. \textit{Strategic Goal:} Wait and see which way the wind blows, ready to join the victor or carve out a neutral buffer state.
\end{itemize}

\subsubsection{Strategic Timers}
\begin{itemize}
\item \textbf{Pass Integrity [8]:} Measures the condition of the fortifications, walls, and defensive works along the pass. Increases after each successful enemy siege or assault that is repelled; decreases when a section of the wall is breached, a gate is burned, or a siege engine establishes a foothold. At 4 a section of the pass is lost (imposing a –1 Position penalty to all defensive actions in that sector); at 8 the pass is compromised and the defenders must fall back to a secondary line.
\item \textbf{Khanate Cohesion [6]:} Tracks the invader's internal unity. Increases when a battle is won, plunder is distributed fairly, or the warlord delivers on promises; decreases when a clan feels slighted, supply runs short, or the warlord's judgement is questioned. At 0 the Khanate fractures into feuding warbands; at 6 it operates with a +1 Position bonus to all rolls.
\item \textbf{Alliance Cohesion [6]:} Gauges the loyalty and morale of the Aeler's lowland allies and mercenary contracts. Increases when the defenders win a decisive victory or hold a key sector of the pass; decreases when the defenders suffer a major loss, a mercenary contract expires unpaid, or a human lord feels abandoned. At 0 a key ally defects or abandons the cause, potentially opening the pass to a flanking maneuver.
\end{itemize}

\subsubsection{Campaign Arcs}
\begin{enumerate}
\item \textbf{The First Wave:} The Khanate's initial assault tests the defenders' strength. It is a swarm of light horse-archers and conscripted infantry, meant to soften the defenses, probe for weaknesses, and gauge the enemy's resolve. \textit{Key Decision:}
\begin{itemize}
\item \textit{Commit Elite Units:} Throw the dwarven stone-warriors and veteran Aeler guards into the breach to hold the line at all costs (risking the loss of irreplaceable troops but gaining a +2 Position bonus to the defense). Success may halt the wave; failure sees the elite units suffer heavy Harm and the defenders' morale shaken.
\item \textit{Use Mercenaries as Cannon-Fodder:} Deploy the hired human levies and dwarven mercenaries as a sacrificial screen, preserving the core Aeler veterans for later (saves veterans but risks a morale penalty if the screen breaks and the mercenaries rout). Success buys time; failure may see the screen collapse and the elite units forced into a desperate last stand.
\end{itemize}
\item \textbf{The Siege of the Gate:} The Khanate brings forward its massive siege towers, rams, and sappers, attempting to batter the main gate of the pass into submission. This becomes a pitched battle of fire, iron, and stone where engineers and heroes decide the day. \textit{Key Decision:}
\begin{itemize}
\item \textit{Focus on Holding the Gate:} Concentrate all available troops and engineers on reinforcing the gate, repairing damage, and pouring boiling oil on attackers (hopes the Khanate's strength will break on the stone; success grants a +1 Position bonus to the defense and slows the siege timer). Risk is that a prolonged siege drains supplies and morale.
\item \textit{Desperate Sortie to Destroy Siege Engines:} Launch a bold counter-attack with shock troops and sappers aimed at burning the siege towers or collapsing the ramps before they can reach the gate (requires a successful Wits+Tactics roll to time the sortie; success destroys a siege engine and advances the Pass Integrity timer by 2; failure sees the sally repulsed with heavy losses and may invite a devastating counter-charge).
\end{itemize}
\item \textbf{The Winter March:} Realizing it will not break the pass by direct assault, the Khanate begins a treacherous flanking march along an ancient, seemingly impassable goat-track that winds through the high cliffs. They have found the one weakness in the Aeler's defenses—a narrow ledge that, if held, could allow them to pour down behind the defenders. \textit{Key Decision:}
\begin{itemize}
\item \textit{Abandon the Pass and Fight a Delaying Action in the Rear:} Pull the main defending force back to a secondary line, using the terrain to channel the Khanate into a killing zone where their numbers cannot be brought to bear (requires a successful Tactics roll to set up the ambush; success gives the defenders Dominant Position on the engagement and may halt the flanking maneuver). Sacrifice is abandoning the primary gate, but it may save the army.
\item \textit{Trust the Track's Treachery:} Hold the pass and rely on the legend that the goat-track is as deadly as the stories say, hoping the Khanate's flank will falter on the icy ledges or be ambushed by hidden Aeler patrols (no special roll needed; maintains the current line but risks the flank succeeding if the legends prove false). Success preserves the pass; failure sees the defenders caught between two hammer blows and the gate lost.
\end{itemize}
\end{enumerate}

\subsubsection{Potential Endings}
\begin{itemize}
\item \textbf{Victory:} The pass holds. The Khanate's coalition is broken, its supplies exhausted, its morale shattered by the relentless stone and cold. The Mountain Kingdom stands, a testament to dwarven stubbornness and Aeler resolve. The king's name becomes a byword for defiance, and his line is remembered in stone-carved sagas.
\item \textbf{Broken Alliance:} The pass is held, but the cost is the dissolution of the alliance. The lowland human allies, having borne the brunt of the fighting and suffered the highest casualties, feel betrayed by Aeler tactical decisions that favored the stone-warriors. The Kingdom is saved, but its friendships are shattered, and its future is more isolated than ever.
\item \textbf{The Mountain Falls:} The Khanate's flanking maneuver succeeds. Their shock troops pour down the goat-track, striking the defenders' rear while the main force presses the gate. The defenders are caught between two hammer blows, the pass is lost, and the Mountain Kingdom is overrun. Survivors are forced into exile, its king a legend of a fallen age. But the Khanate does not burn and pillage—it demands tribute, opens the pass to trade, and establishes a hybrid steppe-settler kingdom, its warlord crowned as a new kind of king.
\end{itemize}

\subsection{The Iron Compact}

\textbf{Premise:} A generation of drought and overpopulation has pushed the Eastern Ykrul clans westward. Their herds are dying, their young warriors restless. A visionary khatun has forged a fragile coalition of clan-lords, promising not plunder but land—fertile valleys and wintering grounds abandoned by a lowland empire that has grown soft and complacent. She does not seek to destroy; she seeks to settle. Her army is a migration, a nation on the move, and its target is the breadbasket of a decadent kingdom. The campaign is a war of movement and negotiation, where supply lines, diplomacy, and the consent of the conquered matter as much as the clash of steel.

\subsubsection{Factions & Actors}
\begin{itemize}
\item \textbf{The Iron Compact:} The Ykrul host—herdsmen, warriors, shamans, and their families—all bound by personal oath to the khatun. Their strength lies in mobility, intimate knowledge of the steppe, and the ability to adapt. \textit{Strategic Goal:} Secure new grazing lands and wintering grounds for the clans; establish a permanent foothold in the lowlands; integrate with the local population rather than displace them.
\item \textbf{The Decadent Kingdom:} A once-great river kingdom, now riven by court faction and complacent in its wealth. Its armies are professional but unmotivated; its fortifications are neglected; its nobles are more interested in political games than defense. \textit{Strategic Goal:} Protect the heartland and the capital, but its leaders are divided on whether to fight, negotiate, or flee.
\item \textbf{The Frontier Lords:} The local nobility who hold the border marches and the contested valleys. They are pragmatic and self-interested; they know the empire's court will not save them. \textit{Strategic Goal:} Survive the Ykrul advance, ideally by cutting a deal that preserves their lands and titles under new suzerainty.
\item \textbf{The Gold Cults:} Merchants and guild-masters who see the Ykrul advance as an opportunity to break the kingdom's trade monopolies. They are willing to fund the khatun's campaign in exchange for favorable trading rights. \textit{Strategic Goal:} Profit from the chaos and secure a place in the new order.
\end{itemize}

\subsubsection{Strategic Timers}
\begin{itemize}
\item \textbf{Compact Unity [8]:} Measures the loyalty of the various clans to the khatun's vision. Increases after each successful conquest, fair distribution of plunder, or victory in battle; decreases when a clan feels slighted, supply runs short, or the khatun's vision is questioned. At 4 a powerful clan may break away; at 8 the Compact is an unstoppable force, operating with a +1 Position bonus to all rolls.
\item \textbf{Kingdom Resolve [6]:} Reflects the lowland kingdom's will to resist. Increases when a city holds out, a relief force is raised, or a noble rallies the court; decreases when a city falls, a noble defects, or the court descends into paralysis. At 0 the kingdom's leadership collapses; at 6 it gains a +1 Position bonus on all defensive actions.
\item \textbf{Frontier Loyalty [6]:} Tracks the allegiance of the border lords and frontier garrisons. Increases when the kingdom sends aid or reinforcements; decreases when aid fails to arrive or the Ykrul offer better terms. At 0 the frontier lords swear fealty to the khatun, opening the path to the heartland.
\end{itemize}

\subsubsection{Campaign Arcs}
\begin{enumerate}
\item \textbf{The Crossing of the Steppe:} The Iron Compact begins its migration westward. They must cross a vast expanse of contested steppe, outrunning or outmaneuvering patrols while gathering supplies and intelligence. \textit{Key Decision:}
\begin{itemize}
\item \textit{Fast and Light:} Move quickly, leaving vulnerable families and herds behind with a rearguard (risks family members being captured; gain +1 Speed for the army). Success reaches the frontier before the kingdom can muster; failure sees the rearguard destroyed and morale shaken.
\item \textit{Safe and Steady:} Keep the Compact together, protecting families and herds at the cost of speed (slower advance; gain +1 Cohesion for the Compact). Success keeps the army intact and ready for the coming fights; failure allows the kingdom to fortify the frontier.
\end{itemize}
\item \textbf{The Siege of the Border Fortresses:} The Ykrul arrive at the first line of frontier fortifications—a chain of fortified towns that control the valleys. They must take these strongholds to secure their wintering grounds. \textit{Key Decision:}
\begin{itemize}
\item \textit{Starve Them Out:} Bypass the forts and cut their supply lines, forcing the garrisons to surrender or desert (requires a successful Logistics roll to maintain the siege; success secures the forts without costly assault; failure allows a relief force to arrive).
\item \textit{Diplomatic Overture:} Send envoys to the frontier lords, offering generous terms: keep their lands and titles, pay a modest tribute, and accept Ykrul suzerainty (requires a successful Presence+Sway roll; success wins the lords' allegiance, gaining a +2 Position bonus for the next phase; failure hardens their resistance).
\end{itemize}
\item \textbf{The Battle of the Ford:} The kingdom's main army arrives to relieve the siege. The Ykrul must defeat this host on the field of battle, using their mobility and tactics to offset the enemy's numbers and armor. \textit{Key Decision:}
\begin{itemize}
\item \textit{The Feigned Retreat:} Pretend to break, draw the kingdom's heavy infantry into a trap where horse-archers and light cavalry can harass them to exhaustion (requires a successful Deception roll; success gives the Ykrul Dominant Position; failure sees the feint turn into a real rout).
\item \textit{The Decisive Charge:} Commit the elite riders to a crushing charge against the kingdom's center, aiming to capture the enemy general and break their morale (requires a successful Tactics roll; success shatters the kingdom's army; failure inflicts heavy casualties on the Ykrul elite).
\end{itemize}
\end{enumerate}

\subsubsection{Potential Endings}
\begin{itemize}
\item \textbf{The New Kingdom:} The Ykrul conquer the river valley, establishing a hybrid steppe-settler state. The khatun becomes the first queen of a new dynasty, ruling with the consent of both her clans and the integrated lowland population. The border lords become her vassals; the Gold Cults become her financiers. The kingdom is transformed, not destroyed.
\item \textbf{The Compromise:} The war ends in a negotiated settlement. The Ykrul receive the contested frontier valleys as a semi-autonomous territory, paying nominal tribute to the lowland crown while maintaining their own laws and customs. The khatun becomes a recognized march-warden, securing the border for the kingdom while preserving her people's way of life.
\item \textbf{The Fracture:} The Ykrul coalition breaks apart before the conquest is complete. A powerful clan-lord, jealous of the khatun's authority, leads a rebellion. The Ykrul are caught between internal strife and the kingdom's counter-attack, and the migration dissolves into a desperate scattering of clans, each fighting for survival.
\end{itemize}

\subsection{The Peacock Throne's Shadow}

\textbf{Premise:} In the heart of the Dhaharan basin, the ancient Peacock Throne—a symbol of millennia-old legitimacy—stands empty. The monarch died without a clear heir, and three rival claimants, each backed by a different power bloc, tear the kingdom apart. The conflict is a multi-front political war where battles are fought for hearts and minds as much as for fields and forts, and where propaganda, oaths, and secret pacts shift as swiftly as the tides.

\subsubsection{Factions \& Actors}
\begin{itemize}
    \item \textbf{Claimant A (The True Heir):} The old king's eldest son, a stern upholder of tradition, backed by the conservative nobility and the Covenant of Mykkiel (a priestly order that guards the ancient rites). \textit{Strategic Goal:} Claim the throne by right of blood and law, restoring the old order and the priest-council's authority.
    \item \textbf{Claimant B (The Reformer):} A younger, charismatic prince, a lover of novelty and progress, supported by the merchant guilds of the river ports and the Temple of Light (a sect that worships the ever-burning flame as a symbol of renewal). \textit{Strategic Goal:} Claim the throne by the will of the people and promise a new era of trade, innovation, and religious tolerance.
    \item \textbf{Claimant C (The Foreign Puppet):} A prince installed by the Veiled Sisters of Ashaan, who see the civil war as a chance to turn their ancient rival into a client state. \textit{Strategic Goal:} Make the kingdom a vassal of the Veiled Throne, securing tribute, military access, and a buffer against the steppe nomads.
\end{itemize}

\subsubsection{Strategic Timers}
\begin{itemize}
    \item \textbf{Kingdom Stability [8]:} Measures the general peace, prosperity, and social cohesion of the Dhaharan people. Increases when a sack or atrocity is avoided, markets remain open, and the common folk feel secure; decreases when a city is sacked, a massacre occurs, or a claimant's troops requisition food brutally. At 0 the Kingdom collapses into anarchy; at 8 it is unified under a strong ruler who enjoys a +1 Position bonus to all rolls.
    \item \textbf{Foreign Intervention [6]:} Gauges the degree to which outside powers are involved in the war. Increases when a claimant's patron dispatches overt military support (troops, siege engines, or naval squadrons); decreases when a patron's interest wanes or their support is withdrawn. At 6 the war becomes a proxy war between major powers, and both sides suffer a -1 Position penalty to all rolls due to divided loyalties and uncertain supply.
    \item \textbf{Peasant Revolt [6]:} Reflects the anger of the common people, who bear the brunt of requisitions, forced conscription, and battlefield devastation. Increases with each army's seizure of grain, each forced levy, or each scorched-earth tactic; decreases when a claimant successfully addresses grievances (e.g., by fixing tax rates or granting land). At 6 the peasants rise up against \textit{all} claimants, throwing the war into chaos and imposing a -2 Position penalty to all military units as loyalty evaporates.
\end{itemize}

\subsubsection{Campaign Arcs}
\begin{enumerate}
    \item \textbf{The First Blood:} The war begins with a single, shocking act—a claimant is assassinated in the night, or a claimant tries to eliminate a rival with a poisoned chalice. The kingdom erupts into a chaotic period of shifting alliances, sudden betrayals, and localized skirmishes as each faction scrambles for advantage. \textit{Key Decision:} The players must decide which claimant to support, and must survive the first few weeks of the civil war—a time when loyalty is fluid and every street corner may hide an assassin.
    \item \textbf{The Battle of the Three Armies:} The three claimants, each having gathered their forces, meet on a single vast plain outside the capital. This is the decisive engagement of the war; one claimant can seize the throne by force and dictate the terms of any peace. \textit{Key Decision:} The players must decide how to tilt the balance of power: they could join the battle on one side (adding their strength to that side's roll), or they could attempt to broker a three-way parley (requiring a successful Presence+Sway roll against a DV that rises with each claimant's suspicion; success grants a temporary cease-fire and the chance to negotiate a power-sharing agreement; failure may see the talks break and the battle proceed with heightened hostility).
    \item \textbf{The Siege of the Capital:} The war is over for all but one claimant, who has retreated to the ancient capital, a maze of stone-courts, bronze-gates, and hidden archives. A siege is laid by the two victors. The battle is not for the city's stones, but for the soul of the kingdom—its legitimacy, its memory, and its future. \textit{Key Decision:} The players must decide how to treat the defeated claimant:
    \begin{itemize}
        \item \textit{Justice:} Put the claimant on trial for treason or tyranny (requires a successful Wits+Law roll; success may see the claimant executed or imprisoned, removing a future threat but potentially creating a martyr).
        \item \textit{Exile:} Strip the claimant of title and lands, sending them into the steppe or the far islands with a modest retinue (removes the immediate danger but leaves a possible future rebel leader in the wilderness).
        \item \textit{Puppet Regime:} Allow the claimant to remain as a figurehead while real power rests with the victorious coalition (requires a successful Presence+Sway roll to maintain the façade; grants the victors stable control but risks future insurgency if the puppet gains popular support).
    \end{itemize}
\end{enumerate}

\subsubsection{Potential Endings}
\begin{itemize}
    \item \textbf{The True Heir's Victory:} The kingdom is unified under a traditionalist, conservative ruler. The Covenant of Mykkiel is strengthened, its rites observed in every town, but the Temple of Light and the merchant guilds are suppressed, their innovations stifled. The kingdom is stable, but ossified, resistant to change, and prone to stagnation.
    \item \textbf{The Reformer's Victory:} The kingdom is unified under a reformist ruler, a monarch of the people who champions the Temple of Light and grants the merchant guilds a seat at the council. The kingdom becomes dynamic and open to new ideas, trade flourishes, but some ancient traditions are lost, and the Covenant's influence wanes.
    \item \textbf{The Puppet's Victory:} The kingdom is nominally unified, but is now a client state of the Veiled Throne. It enjoys wealth and peace, yet its strings are pulled from Ashaan; its future is bright, but it is not their own, and any move toward independence risks swift retaliation.
    \item \textbf{The Peasant's Victory:} Weary of war, the common folk rise up, depose all claimants, and declare a fragile Republic. The Kingdom is free of monarchs, but the new government is weak, fractious, and vulnerable to invasion from the steppe or the sea; its survival depends on the ability of the citizens to maintain unity in the face of external threat.
\end{itemize}

\subsection{Optional Twists (Applicable to Any Seed)}
\begin{itemize}
    \item \textbf{The Treacherous Ally:} A trusted ally—a major lord, mercenary captain, or influential guildmaster—is secretly in the pay of the enemy. Their betrayal comes at a crucial moment (e.g., during a siege, before a decisive battle, or while negotiating a treaty), shifting the balance of the campaign and forcing the PCs to reevaluate their loyalties.
    \item \textbf{The Divine Intervention:} A patron spirit, an elemental force, or a forgotten god manifests at a critical juncture—perhaps appearing on the battlefield as a storm, a sudden fog, or a vision in the commander's dream. Victory now depends on faith and ritual as much as steel; those who perform the proper rites may gain a boon (e.g., +1 Position, a free reroll, or temporary immunity to morale checks), while those who ignore the sign suffer a penalty.
    \item \textbf{The Forgotten Enemy:} The focus of the war shifts when a previously dormant threat—an ancient dragon slumbering beneath a mountain, a virulent plague spreading from the marshes, or a rival empire stirring beyond the horizon—awakens or attacks. Both sides must make a terrible choice: unite against the new peril or risk annihilation. This twist can trigger a temporary alliance, a desperate peace, or a frenzied scramble for survival.
\end{itemize}

\section{Epilogue: The General's Final Counsel}

\textit{``I have given you the tools. I have given you the principles. I have given you the warning that war is a fire that burns those who touch it—and those who only stand near it. What you do with these tools, these principles, this knowledge, is your choice. But remember: war is not a game. It is a tragedy. If you must play it, play it well—and try to leave something worth saving behind.''}

\textit{``The road is long. The war is hard. The scars will remain. But if you fight for something worth fighting for—something that can survive the war and grow after it—then the scars will be a testimony to what you built, not a monument to what you burned.''}

\textit{``Go now. The campaign awaits.''}

\noindent— General Marcus Valerius, Eastern Marches Command, Ecktoria

\end{document}