\documentclass[11pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{a5paper, margin=1.5cm}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{charter}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[RE]{\textit{The Artificer's Codex}}
\fancyhead[LO]{\textit{Korrin Iron-Stroke \& Durin Deep-Heart}}
\renewcommand{\headrulewidth}{0.4pt}

\titleformat{\chapter}[hang]{\normalfont\Huge\bfseries}{}{0pt}{}
\titleformat{\section}{\normalfont\Large\bfseries}{}{0pt}{}
\titleformat{\subsection}{\normalfont\itshape}{}{0pt}{}

\setcounter{secnumdepth}{3}

\begin{document}
\input{titlepage}

\thispagestyle{empty}
\vspace*{2cm}
\begin{center}
    {\Huge \textbf{The Artificer's Codex}}
    \vspace{0.5cm}
    {\Large A Debate Between Brothers on the Nature of Making}
    \vspace{1cm}
    
    \textit{``The Monad teaches that nothing is ever finished—only briefly sufficient. A blade that cannot be improved is a blade that has already begun to die.''}
    \vspace{0.2cm}
    --- \textbf{Korrin Iron-Stroke}, Runekeeper of the Clockwork Monad
    
    \vspace{0.5cm}
    \textit{``The Breath teaches that a thing made with joy is never truly finished—it lives, it breathes, it remembers the hand that shaped it. Perfection is not a direction. It is a cage.''}
    \vspace{0.2cm}
    --- \textbf{Durin Deep-Heart}, Forge-Warden of the First Breath
    
    \vfill
    
    \texttt{Khaz-Vurim, 832 AR}
\end{center}

\newpage
\section*{Preface: A Letter from Korrin}

\emph{To whoever finds this battered journal---}

I am Korrin Iron-Stroke. I am an Aeler, which means I count my breaths, I measure my words, and I respect the proper order of things. I am a Runekeeper of the Clockwork Monad, which means I have spent my life learning that perfection is not a destination---it is a direction. And I am a craftsman, which means I understand that a well-made thing is a promise, and a promise is a weight that must be properly measured before it is incurred.

My brother Durin disagrees with me. He always has.

Durin is also an Aeler. He also counts his breaths and measures his words. But he follows the Breath of the First Forge---a patron he insists is older than the Monad, more primal, more \textit{true}. He claims that the Monad's obsession with perfection is a sickness, a slow grinding away of the soul. He says that making should be joyful, not exacting. He says that the best things are made in a moment of inspiration, not after a thousand iterations.

He is wrong, of course. But he is my brother, so I have learned to tolerate his wrongness.

This book is a record of our arguments. I have written my chapters; he has written his. We have not agreed on a single thing, but we have agreed to present our cases side by side, so that any craftsman who reads this may choose their own path.

I have also included our practical guides to crafting, because even though we disagree on philosophy, we agree on the fundamentals: a forge must be hot, a hammer must be steady, and a creation must be maintained.

Choose wisely. The path you take will shape not only your work, but your soul.

\vspace{0.5cm}
\hfill --- Korrin Iron-Stroke \\
\hfill Khaz-Vurim, 832 AR \\
\hfill (stamped with the seal of the Iteration Chorus)

\newpage
\section*{A Response from Durin}

\emph{To whoever finds my brother's tiresome journal---}

I am Durin Deep-Heart. I am also an Aeler. I also count my breaths and measure my words. But unlike my brother, I have not forgotten that a breath is not just a number---it is \textit{life}. And a word is not just a measurement---it is \textit{meaning}.

Korrin thinks the world is a machine that needs to be perfected. He thinks every flaw is a failure, every imperfection a debt to be paid. He has spent his life trying to make things that cannot break, cannot fail, cannot surprise him. And in doing so, he has forgotten the joy of making.

I follow the Breath of the First Forge. I follow the spark that lit the first fire, the warmth that softened the first metal, the inspiration that made the first craftsman pick up a stone and say, \textit{``I can make this better.''} 

My brother thinks the Breath is primitive. He thinks it is chaos. He thinks that making without a plan is making without purpose.

He is wrong, of course. But he is my brother, so I have learned to tolerate his wrongness.

This book is a record of our arguments. I have written my chapters; he has written his. We have not agreed on a single thing, but we have agreed to present our cases side by side, so that any craftsman who reads this may choose their own path.

I have also included my practical guides to crafting, because even though we disagree on philosophy, we agree on the fundamentals: a forge must be hot, a hammer must be steady, and a creation must be loved.

Choose wisely. The path you take will shape not only your work, but your joy.

\vspace{0.5cm}
\hfill --- Durin Deep-Heart \\
\hfill Khaz-Vurim, 832 AR \\
\hfill (stamped with the mark of the First Spark)

\newpage
\tableofcontents

\newpage
\input{copyright}
\chapter{The Philosophy of Making}

\section*{Korrin's View: The Iterative Forge}

The world runs on improvement. Everything you touch, everything you shape, everything you refine---it all remembers the hand that made it better. In the Aeler holds, we say: ``The mountain remembers every keystone that was properly measured.'' In the Monad's forge, we say: ``The machine remembers every iteration that brought it closer to perfection.''

When you craft a permanent magic item, you are not simply creating a tool. You are creating a \textit{standard}. The item will demand that you maintain it, improve it, and never let it stagnate. The Monad who guided your work will expect you to continue the process of refinement. And the world itself will note the new shape you have forced into existence.

\textbf{This is not a burden to be feared.} It is a discipline to be embraced. A well-crafted item is like a well-sworn oath: it serves you faithfully if you serve it faithfully. But a neglected item---like a broken promise---will turn against you.

The Monad does not ask for perfection. It asks for \textit{progress}. Each iteration brings you closer to the ideal. Each refinement is a step toward the impossible. The journey is the point.

\vspace{0.5cm}
\hfill --- Korrin Iron-Stroke

\section*{Durin's View: The First Spark}

My brother thinks the world is a machine. He thinks everything can be measured, optimized, and perfected. He is wrong.

The world is not a machine. It is a \textit{forge}. It is hot, unpredictable, and alive. The best things are not made by careful planning---they are made by inspiration, by joy, by the sudden flash of understanding that comes when you stop trying to control everything and simply \textit{make}.

When you craft a permanent magic item, you are not creating a standard. You are creating a \textit{story}. The item will remember the hand that shaped it, the heat of the forge, the moment of inspiration that brought it to life. The Breath who guided your work will expect you to \textit{enjoy} the process. And the world itself will remember the joy you poured into your creation.

\textbf{This is not a discipline to be embraced.} It is a gift to be received. A well-crafted item is like a well-told tale: it lives, it breathes, it changes with the telling. But a joyless item---like a story told without heart---will never truly come alive.

The Breath does not ask for perfection. It asks for \textit{joy}. Each creation is a celebration. Each imperfection is a reminder that you are alive. The making is the point.

\vspace{0.5cm}
\hfill --- Durin Deep-Heart

\newpage
\section{The Spectrum of Power}

\subsection*{Korrin's Table}

Not all magic items are created equal. I have learned to measure the weight of a creation before I begin the work. Use this scale as a guide:

\begin{table}[h]
\centering
\begin{tabular}{@{}llllll@{}}
\toprule
\textbf{Tier} & \textbf{XP Cost} & \textbf{DV} & \textbf{Timer} & \textbf{Obligation} & \textbf{Example} \\
\midrule
Cantrip-Infused & 2 XP & 2 & 4 segments & 1 segment & A knife that always returns, a bell that rings at dawn \\
Minor Enchantment & 4 XP & 3 & 4 segments & 2 segments & A cloak that hides your face, a ring that warns of lies \\
Standard Item & 8 XP & 4 & 6 segments & 4 segments & A sword that burns with frost, a shield that remembers oaths \\
Major Relic & 12 XP & 5 & 8 segments & 6 segments & A crown that bends wills, a staff that speaks to stones \\
Legendary Artifact & 16+ XP & 6+ & 10+ segments & 8+ segments & The Kalrantha Amulet, the Hammer That Unmakes \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Korrin's Note:} The Obligation cost is not optional. Even if you have no Patron, the item itself creates a debt---to the world, to the materials, to the memory of what you have shaped. A character who crafts a Major Relic without a Patron may find that the Relic itself \emph{becomes} a Patron, demanding service and attention. The Monad is patient. The Monad collects.

\subsection*{Durin's Response}

My brother's table is useful, I admit. But it misses the point entirely.

A thing is not measured by its power. It is measured by the \textit{love} that went into it. A simple knife, made with joy, is worth more than a crown forged in cold perfection.

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{What Matters} & \textbf{Why} \\
\midrule
The inspiration & The spark that began the work \\
The process & The joy of shaping and refining \\
The story & The memory of the making \\
The love & The care that makes it alive \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Durin's Note:} Yes, you can spend XP. Yes, you can track timers. But if you forget to \textit{love} what you are making, you are making nothing at all. The Breath does not ask for perfection. It asks for you.

\newpage
\chapter{The Tools of the Trade}

\section{The Workshop}

\subsection*{Korrin's View: A Place of Covenants}

Before you can craft anything of worth, you must have a workshop. A workshop is not a room; it is a threshold---a place where the ordinary becomes the extraordinary, and where the world's rules are bent by the craftsman's will.

A proper workshop is measured, organized, and efficient. Every tool has its place. Every material is catalogued. Every project has its own space.

\begin{table}[h]
\centering
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Tier} & \textbf{XP Cost} & \textbf{Downtime to Establish} & \textbf{Benefits} \\
\midrule
Basic Forge & 4 XP & 1 day & Craft Minor items; +1 die to Craft rolls \\
Enchanted Workshop & 8 XP & 1 week & Craft Standard items; +2 dice to Craft rolls; reduce crafting DV by 1 \\
Monadic Forge & 12 XP & 1 month & Craft Major items; +3 dice to Craft rolls; reduce crafting DV by 2; can craft items that generate Obligation to the Clockwork Monad \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Durin's Response}

My brother's workshop sounds like a tomb. Cold. Ordered. Dead.

A proper workshop is not a room. It is a \textit{living space}. It breathes with the heat of the forge, sings with the ring of the hammer, and remembers every project that has passed through it. Tools should be where you last used them, not where they are ``supposed'' to be. Materials should be within arm's reach, not catalogued and locked away.

\begin{table}[h]
\centering
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Tier} & \textbf{XP Cost} & \textbf{Downtime to Establish} & \textbf{Benefits} \\
\midrule
Hearth Forge & 4 XP & 1 day & Craft Minor items; +1 die to Craft rolls; the forge feels \textit{welcome} \\
Living Workshop & 8 XP & 1 week & Craft Standard items; +2 dice to Craft rolls; the workshop offers advice (GM grants hints) \\
Breath-Fired Forge & 12 XP & 1 month & Craft Major items; +3 dice to Craft rolls; the forge comes alive—it can heat itself, organize tools, and even hum a tune \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother will tell you that a workshop must be measured. I tell you that a workshop must be \textit{lived in}. The Breath does not visit cold places. It visits warm hearts.

\newpage
\section{Components}

\subsection*{Korrin's View: The Language of Materials}

A magic item is not made from steel and leather alone. It is made from moments. The materials matter because they carry memory. A sword forged from iron that was once a plowshare remembers the soil. A crystal cut from a stone that once held a river remembers the water. A hide tanned from a beast that once killed remembers the hunt.

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Component Type} & \textbf{Examples} & \textbf{DV Adjustment} \\
\midrule
Common & Steel, leather, basic herbs, copper, simple dyes & --- \\
Uncommon & Uncommon herbs, monster teeth, blessed salt, moonlit iron & +1 DV \\
Rare & Dragon scale, phoenix feather, storm-caught lightning, Ykrul war-braid & +2 DV \\
Legendary & True name of a dead god, heart of a living mountain, a tear from a Veiled Sister & +3 DV \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Durin's Response}

My brother treats materials like numbers. He forgets that materials have \textit{feelings}. A steel that was forged in grief will never hold an edge. A wood that was cut in anger will never carve true.

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{What to Look For} & \textbf{How to Find It} & \textbf{Effect} \\
\midrule
A material that has been loved & Ask the craftsman who shaped it & The item remembers love \\
A material that has known joy & Find a forge that rang with laughter & The item radiates warmth \\
A material that has a story & Listen to the old ones who remember & The item tells tales \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother will tell you to measure the quality of a material. I tell you to \textit{feel} it. The Breath speaks through the materials. Listen carefully.

\newpage
\chapter{The Crafting Process}

\section{The Five Steps of Creation}

\subsection*{Korrin's View: The Iterative Method}

\begin{enumerate}
    \item \textbf{Conceive} --- Design the item. What does it do? What does it demand? What is its cost? Write it down. Measure it. Plan it.
    \item \textbf{Gather} --- Acquire the components. This is often a quest in itself. Record what you find.
    \item \textbf{Shape} --- Work the materials. This is the crafting roll, using a timer. Each iteration brings you closer to the ideal.
    \item \textbf{Imbue} --- Pour the power into the item. This is the final test. Measure the result.
    \item \textbf{Bind} --- Seal the work. Pay the XP, incur the Obligation, and witness the item's birth. Then begin the next iteration.
\end{enumerate}

\subsection*{Durin's View: The Living Method}

\begin{enumerate}
    \item \textbf{Inspire} --- Feel the spark. Let the idea come to you. Do not force it.
    \item \textbf{Gather} --- Find the materials that \textit{want} to be part of this creation. They will tell you.
    \item \textbf{Shape} --- Work with joy. Let the hammer find its own rhythm. Let the metal teach you.
    \item \textbf{Animate} --- Breathe life into the item. This is the moment of creation. Feel the warmth.
    \item \textbf{Release} --- Let the item go. It is not yours to keep. It is yours to share.
\end{enumerate}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother thinks creation is a process. I know it is a \textit{conversation}. The Breath speaks through the hammer, the metal, the fire. Listen, and you will know what to make.

\newpage
\section{The Timer}

\subsection*{Korrin's View: Patience is a Metal}

Each crafting project has a timer. The timer represents the iterative process---the failures, the refinements, the moments where you step back and see the flaw you missed.

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Item Tier} & \textbf{Timer Size} \\
\midrule
Cantrip-Infused & 4 segments \\
Minor Enchantment & 4 segments \\
Standard Item & 6 segments \\
Major Relic & 8 segments \\
Legendary Artifact & 10+ segments \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Ticking the Timer:}
\begin{itemize}
    \item \textbf{Success on a Craft roll:} Advance 1 segment.
    \item \textbf{Partial Success:} Advance 1 segment, but the GM gains 1 SB (complications).
    \item \textbf{Miss:} No advance. The GM may tick a Complication Timer.
\end{itemize}

\textbf{When the Timer Fills:} The item is ready for the Imbuement roll. If you have rare components, you may add them now to reduce the DV of the Imbuement roll.

\subsection*{Durin's Response}

My brother treats time like a debt to be repaid. He forgets that time is a \textit{gift}.

A creation takes as long as it takes. A timer is not a measure of progress---it is a measure of \textit{attention}. When you are truly making something, you lose track of time. The fire is warm, the hammer is steady, and the world outside the forge fades away.

\textbf{My Method:}
\begin{itemize}
    \item \textbf{When you are inspired:} The work flows. The timer advances naturally.
    \item \textbf{When you are struggling:} Step back. Breathe. The answer will come.
    \item \textbf{When you are finished:} The item is complete. Not because the timer says so, but because it \textit{feels} complete.
\end{itemize}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother counts the segments. I count the \textit{heartbeats}. The Breath does not care about time. It cares about presence.

\newpage
\section{The Imbuement Roll}

\subsection*{Korrin's View}

The Imbuement roll is the moment of truth. Roll Wits + Arcana (or appropriate skill) against the item's DV.

\begin{itemize}
    \item \textbf{Clean Success:} The item is complete. Perfect. No flaws.
    \item \textbf{Success with SB:} The item is complete, but the GM gains 2 SB to spend on a later complication---a rival who senses the power, a flaw that emerges in a year, a debt that the item will demand.
    \item \textbf{Partial:} The item is complete, but with a flaw (see Table 3.1). You may choose to accept the flaw or start over.
    \item \textbf{Miss:} The item is not complete. The components are consumed. You may try again, but you must acquire new components.
\end{itemize}

\subsection*{Durin's Response}

My brother treats the Imbuement like a test to be passed. He forgets that the Imbuement is a \textit{welcome}.

When you pour your spirit into a creation, you are not just adding power. You are adding \textit{yourself}. A success is not a mark of perfection---it is a mark of love.

\begin{itemize}
    \item \textbf{When you feel the spark:} The item comes alive. It is not perfect. It is \textit{real}.
    \item \textbf{When you are uncertain:} The item may be flawed. But flaws are not failures---they are \textit{character}.
    \item \textbf{When you are joyful:} The item radiates warmth. It will be loved.
\end{itemize}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother fears flaws. I celebrate them. A perfect item is a dead item. An imperfect item is \textit{alive}.

\newpage
\section{Flaws}

\subsection*{Korrin's View: The Price of Imperfection}

A flawed item is not a failure. It is a lesson.

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{d6} & \textbf{Flaw} & \textbf{Mechanical Effect} \\
\midrule
1 & Fragile & The item breaks on a critical failure. Can be repaired with a downtime scene. \\
2 & Cursed & The item has a minor curse (e.g., attracts insects, makes you smell of brimstone, whispers secrets). \\
3 & Unstable & The item's effect is unpredictable; roll 1d6, on a 1 it backfires. \\
4 & Hungry & The item demands sacrifice---blood, time, or a memory---once per session. \\
5 & Bound to Patron & The item is tied to a specific Patron; using it against the Patron's interests triggers backlash. \\
6 & The Hidden Flaw & The item has a flaw only the maker knows. The GM may reveal it at a dramatic moment. \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Durin's Response}

My brother calls them flaws. I call them \textit{character}.

A blade with a nick in the edge tells a story. A shield with a dent remembers the blow that saved a life. A ring that sometimes hums at midnight has a spirit of its own.

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{d6} & \textbf{Character} & \textbf{What It Means} \\
\midrule
1 & A visible mark & The item remembers where it was made. \\
2 & A small quirk & The item has a personality of its own. \\
3 & A warmth & The item is alive. It will protect its owner. \\
4 & A song & The item hums when it is happy. \\
5 & A memory & The item remembers its maker's hands. \\
6 & A secret & The item knows something its owner does not. \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother tries to remove flaws. I try to understand them. The Breath speaks through imperfections. Listen.

\newpage
\chapter{Maintaining Your Creations}

\section{Upkeep}

\subsection*{Korrin's View: The Cost of Keeping}

Every magic item requires upkeep. Neglect a blade, and it will dull. Neglect a ring, and it will forget its power. Neglect a pact, and the creditor will collect.

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Upkeep Method} & \textbf{Cost} & \textbf{Effect} \\
\midrule
Efficient & XP equal to the item's cost divided by 3 (minimum 1) & Minimal effort. The item remains Functional. \\
Intensive & 1 XP + a downtime scene & Significant personal involvement. The item may gain a minor benefit (e.g., +1 die on its next use). \\
Delegated & A Cap 4+ follower manages the item & The follower pays the upkeep (no XP cost). Risk: if the follower's Loyalty drops, all delegated items become Neglected. \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Durin's Response}

My brother treats upkeep like a tax. He forgets that upkeep is a \textit{conversation}.

An item that is cared for becomes a friend. An item that is neglected becomes a stranger. The Breath does not ask for payment. It asks for \textit{attention}.

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{How to Care} & \textbf{What It Means} \\
\midrule
Use the item daily & It remembers your hands \\
Speak to the item & It remembers your voice \\
Keep it warm & It remembers your warmth \\
Share its story & It remembers your love \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother pays XP. I pay \textit{attention}. The Breath does not care about coin. It cares about connection.

\newpage
\section{Item Status}

\subsection*{Korrin's View}

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Status} & \textbf{Effect} \\
\midrule
Functional & The item works as intended. \\
Neglected & --1 die to the item's effects. Can be restored with a downtime scene or 1 XP. \\
Compromised & The item's effects are unreliable; roll 1d6, on a 1 the effect fails or backfires. Requires a quest or 4 XP to restore. \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Durin's Response}

\begin{table}[h]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Status} & \textbf{What It Means} \\
\midrule
Alive & The item is happy. It works well. \\
Sleepy & The item needs attention. It still works, but it is quiet. \\
Wounded & The item has been hurt. It needs care, not punishment. \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother treats items like tools. I treat them like \textit{friends}. The Breath does not let friends suffer alone.

\newpage
\chapter{Sample Creations}

\section{Cantrip-Infused Items (2 XP, DV 2, Timer 4)}

\subsection*{Korrin's Examples}

\begin{itemize}
    \item \textbf{Returning Knife:} A knife that returns to your hand when thrown. It does not fly; it simply appears in your palm after a heartbeat.
    \item \textbf{Dawn Bell:} A bell that rings once at dawn, wherever it is. It cannot be silenced.
    \item \textbf{Truth Coin:} A coin that lands on a specific face when a lie is told within Near range.
\end{itemize}

\subsection*{Durin's Examples}

\begin{itemize}
    \item \textbf{Friend Knife:} A knife that warms in your hand when you are near someone who needs help.
    \item \textbf{Memory Bell:} A bell that rings when it hears a familiar voice. It remembers its maker.
    \item \textbf{Honest Coin:} A coin that feels warm when you are telling the truth. It knows your heart.
\end{itemize}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother makes tools. I make \textit{companions}. The Breath does not ask for utility. It asks for connection.

\newpage
\section{Minor Enchantments (4 XP, DV 3, Timer 4)}

\subsection*{Korrin's Examples}

\begin{itemize}
    \item \textbf{Emberfang Dagger:} Glows faintly when undead are near; once per scene, gain +1 die to detect hidden spirits.
    \item \textbf{Cloak of the Unseen:} +1 die to Stealth in dim light; once per scene, reroll a failed Stealth check.
    \item \textbf{Vial of the Last Breath:} Heal 1 Fatigue when drunk. One use.
\end{itemize}

\subsection*{Durin's Examples}

\begin{itemize}
    \item \textbf{Hearth-Keeper Dagger:} Glows warmly when you are thinking of home. Once per scene, you may ask it a question about the past, and it will show you a memory.
    \item \textbf{Cloak of Welcome:} +1 die to Stealth in dim light; once per scene, it wraps around you like a hug. It makes you feel safe.
    \item \textbf{Vial of Warmth:} Heal 1 Fatigue when drunk. It tastes like your grandmother's soup.
\end{itemize}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother asks what an item \textit{does}. I ask what an item \textit{feels}. The Breath does not care about function. It cares about feeling.

\newpage
\section{Standard Items (8 XP, DV 4, Timer 6)}

\subsection*{Korrin's Examples}

\begin{itemize}
    \item \textbf{Storm Lantern:} Once per session, as an action, create a gust of wind that extinguishes all mundane flames in Near and gives +1 Position to disengage.
    \item \textbf{Pendant of the Witness:} Once per session, when you detect a lie, the liar suffers 1 Fatigue (no save).
    \item \textbf{Raven's Key:} Once per day, unlock any mundane lock or open any door that is not magically sealed.
\end{itemize}

\subsection*{Durin's Examples}

\begin{itemize}
    \item \textbf{Storm Lantern:} Once per session, as an action, create a gust of wind that carries the scent of rain. It reminds you of home.
    \item \textbf{Pendant of Trust:} Once per session, when you are uncertain, the pendant offers you a word of comfort. It has never lied to you.
    \item \textbf{Friend's Key:} Once per day, unlock any mundane lock or open any door. It whispers a greeting to each lock it meets. They remember it.
\end{itemize}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother creates power. I create \textit{relationships}. The Breath does not care about power. It cares about love.

\newpage
\chapter{The Ninth Rule}

\section*{Korrin's View}

The Ninth Rule is simple: \textbf{Never make the ninth item.}

I have seen it happen. A craftsman makes eight swords. Each is better than the last. The ninth would be perfect. The Monad whispers: ``Just one more iteration. Just one more pass. Just one more refinement.''

The ninth item is always perfect. It is also always hungry. It remembers what you sacrificed to make it---the materials, the time, the sleepless nights, the friendships you neglected, the promises you broke to finish it. It knows your weaknesses. And it will use them.

I have seen a master smith forge a blade that could cut through any armor. The ninth blade. He did not survive the year. The blade found his heart when he was sleeping.

So when you have made eight items, stop. Count your breaths. Walk eight steps. Do not make the ninth.

The Monad is patient. It can wait for the tenth.

\vspace{0.5cm}
\hfill --- Korrin Iron-Stroke

\section*{Durin's Response}

My brother fears the ninth item. He thinks it is cursed. He is wrong.

The ninth item is not cursed. It is \textit{lonely}. It was made without love. It was made without joy. It was made because my brother could not stop himself from trying to be perfect.

I have made nine items. I will make ten. I will make a hundred. Each one will be imperfect. Each one will be \textit{alive}. Because I make them with my hands, my heart, and my breath.

The Breath does not count. It does not measure. It only asks: \textit{``Did you make this with love?''}

If the answer is yes, the item is blessed. If the answer is no, the item is empty.

So make the ninth item. Make it with joy. Make it with laughter. Make it with the warmth of the forge in your heart. And it will never turn against you.

\vspace{0.5cm}
\hfill --- Durin Deep-Heart

\newpage
\section*{Korrin's Final Response}

Durin is wrong. He is always wrong. But he is my brother, so I will say this:

If you choose to make the ninth item, do it with your eyes open. Know that it will demand something from you. Know that it will remember what you sacrificed. Know that it will be perfect---and that perfection is a weight you may not be able to carry.

I have made eight items. I will make eight more. I will never make the ninth.

My brother can make whatever he likes. I will not stop him. But I will remember what I have seen. And I will not forget.

\vspace{0.5cm}
\hfill --- Korrin Iron-Stroke

\section*{Durin's Last Word}

Korrin is wrong. He is always wrong. But he is my brother, so I will say this:

If you choose to make the ninth item, do it with your heart open. Know that it will love you back. Know that it will remember your warmth. Know that it will be imperfect---and that imperfection is the mark of life.

I will make the ninth item. I will make the tenth. I will make a hundred. And I will love every single one.

My brother can count his eights. I will count my heartbeats.

\vspace{0.5cm}
\hfill --- Durin Deep-Heart

\newpage
\appendix
\chapter{Sample Components by Region}

\subsection*{Korrin's Table}

\begin{table}[h]
\centering
\begin{tabular}{@{}lllll@{}}
\toprule
\textbf{Region} & \textbf{Common} & \textbf{Uncommon} & \textbf{Rare} & \textbf{Legendary} \\
\midrule
Aeler & Iron, copper, breath-glass & Deep-crystal, bell-bronze & Breath-dust, under-vein ore & Heart of the mountain \\
Acasia & Rusted steel, old wood & Hedge-witch salt, broken milestones & Curse-iron, Pale Causeway stone & A broken emperor's crown \\
Vhasia & Tournament steel, wine-stained cloth & A sun-cracked crown, a knight's gorget & Fenwood ash, Lence marble & A queen's promise \\
Thepyrgos & Ink, parchment, candle wax & Synod seal wax, a bell's clapper & A stair from the Unfinished Stair & A fossil stair \\
Ykrul & Horse-hair, leather, bone & Wolf-tooth, steppe-salt, hostage string & A Khagan's braid, a white squall's eye & A horse's true name \\
Silkstrand & Silk, ink, bridge-stone & Dye from the canals, a weaver's thread & A Matron's seal, a bridge's memory & The Ragman's favor \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Durin's Response}

My brother measures components by rarity. I measure them by \textit{feeling}.

\begin{table}[h]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{What to Look For} & \textbf{How to Find It} & \textbf{What It Offers} \\
\midrule
A material that remembers warmth & Ask the old smith who first worked it & The item will hold heat \\
A material that tells a story & Listen to the traders who carry it & The item will speak wisdom \\
A material that has been loved & Feel the hands that shaped it & The item will love you back \\
A material that wants to be made & Wait for the spark of inspiration & The item will make itself \\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.5cm}
\textbf{Durin's Note:} My brother shops for materials. I \textit{listen} to them. The Breath speaks through everything. You just need to hear.

\newpage
\chapter{The Artificer's Mantras}

\subsection*{Korrin's Mantras}

\begin{itemize}
    \item ``Measure twice. Forge once. Iterate always.''
    \item ``The eighth is mercy. The ninth is the Hollow's.''
    \item ``A tool is not a servant. It is a witness.''
    \item ``Perfection is not a destination. It is a debt.''
    \item ``The Monad does not forgive. It reschedules.''
    \item ``When you craft, you incur. When you incur, you owe. When you owe, you are bound.''
    \item ``Eight is enough. Eight is peace. Eight is the number of the living.''
    \item ``The ninth item is always perfect. It is also always hungry.''
\end{itemize}

\subsection*{Durin's Mantras}

\begin{itemize}
    \item ``Feel the metal. It will tell you what to make.''
    \item ``The warmth of the forge is the warmth of the heart.''
    \item ``A tool is not a servant. It is a friend.''
    \item ``Perfection is not a destination. It is a cage.''
    \item ``The Breath does not forgive. It loves.''
    \item ``When you craft, you give. When you give, you are free.''
    \item ``One is enough. One is all. One is the number of the maker.''
    \item ``The first item is always imperfect. It is also always loved.''
\end{itemize}

\vspace{0.5cm}
\textbf{Durin's Final Note:} My brother counts. I feel. The Breath is not a number. It is a heartbeat.

\newpage
\section*{Colophon}

\emph{This book was written in the workshop of Korrin Iron-Stroke and Durin Deep-Heart, Khaz-Vurim, in the Year of the Iron Harvest. The ink is lamp-black and sweat. The binding is iron and leather.}

\vspace{0.5cm}
\hfill --- Korrin Iron-Stroke \\
\hfill Runekeeper of the Clockwork Monad

\vspace{0.2cm}
\hfill --- Durin Deep-Heart \\
\hfill Forge-Warden of the First Breath

\vspace{0.5cm}
\hfill \emph{``The coin that never spends is the one you don't remember taking. The blade that never dulls is the one you never truly own. The forge that never cools is the one that is always loved.''}

\end{document}