% Assets & Worldly Patrons — Fate's Edge Module (v0.2)
% Single-file LaTeX source, print-ready
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{array}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{framed}
\hypersetup{colorlinks=true, linkcolor=black, urlcolor=blue, citecolor=black}
\setlist{noitemsep, topsep=4pt, leftmargin=1.2em}
\titleformat{\section}{\normalfont\Large\bfseries}{}{0pt}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{}{0pt}{}
\titleformat{\subsubsection}{\normalfont\normalsize\bfseries}{}{0pt}{}

\begin{document}
\begin{center}
{\LARGE \textbf{Assets \& Worldly Patrons}}\\[3pt]
{\large A Fate's Edge module for holdings, leases, and mortal patronage}\\[2pt]
\textit{Version 0.2 (Playtest Revised)}
\end{center}
\vspace{0.25em}
\hrule\vspace{0.75em}

\section*{Design Goals}
\begin{itemize}
  \item \textbf{Make holdings matter:} Assets have tags, strings, upkeep, and project clocks that change scenes.
  \item \textbf{Patrons with teeth:} Patronage brings Position, seals, and coin---and obligations, audits, and scandals.
  \item \textbf{Use the core:} Position/DV, SB, clocks, Favor/Leverage/Exposure, Reputation, Strings. No new dice math.
  \item \textbf{Low bookkeeping:} Track names, tags, and a few clocks; orders and events move the world.
  \item \textbf{Campaign continuity:} Assets and patrons persist and evolve across story arcs.
\end{itemize}

\section*{Quickstart (2 minutes)}
\begin{enumerate}
  \item Pick/design an \textbf{Asset}; mark \textbf{Integrity [6]} and \textbf{Resources [6]}.
  \item Name a \textbf{Patron} (optional); set \textbf{Patron Tier} and \textbf{Obligations}.
  \item Attach 1--2 \textbf{Strings} (permits, seals, routes, rites).
  \item Choose 1--2 \textbf{Project Clocks [4--8]} to improve or expand the asset.
  \item Each session: issue \textbf{Asset Orders}, confront \textbf{Events \& Audits}, and tick \textbf{Upkeep}.
\end{enumerate}

\tableofcontents
\newpage

\section{Asset Sheet (Template \& Tags)}
\subsection{Asset Sheet (Template)}
\begin{framed}
\noindent\textbf{[ASSET NAME]}\\
\textbf{Type:} safehouse / barge company / workshop / archive / shrine / foundry line / office / permit / theatre / farm / mine / caravan yard / canal lock lease\\
\textbf{Locale:} district \& city (matters for Position \& dials)\\
\textbf{Strings (1--3):} permits, seals, routes, rites, keys\\
\textbf{Tags (2--4):} see tag lists below\\
\textbf{Tracks:}\\
\hspace{10pt}- Integrity [6]: condition/standing; fill $\rightarrow$ shutdown, collapse, or seizure\\
\hspace{10pt}- Resources [6]: cashflow, inventory, staff slack; empty $\rightarrow$ shortages/concessions\\
\hspace{10pt}- Heat [6] (optional): locals' attention; fill $\rightarrow$ protest, inquiry, rough visit\\
\textbf{Projects (0--3):} named [4--8] clocks to upgrade, expand, or pivot\\
\textbf{Upkeep:} coin/favor/rites required each cycle (choose 1--2)\\
\textbf{Legacy Notes:} campaign events, reputation effects, historical significance\\
\textbf{Notes:} staff names, neighbors, rivals, liens
\end{framed}

\subsection{Asset Tags (pick 2--4)}
\begin{itemize}
  \item \textbf{Fortified} (doors, shutters, hidden room) --- \emph{Shield/Petition} here starts \textbf{Controlled}.
  \item \textbf{Sanctified} (rites, hospitality) --- breaking rites here ticks \textbf{Exposure +1} (offender).
  \item \textbf{Licensed} (writ, seal) --- \emph{Broker/Petition} \textbf{DV --1} in scope of license.
  \item \textbf{Secret} (concealed use) --- first \emph{Expose} against you here starts \textbf{Desperate} for the attacker.
  \item \textbf{Crowd-Facing} (shop, theater) --- \emph{Audience tags} created here persist an extra scene.
  \item \textbf{Hazardous} (kilns, reagents) --- on 1s, GM may start \textbf{Accident [4]} instead of other SB spend.
  \item \textbf{Mobile} (barge/caravan) --- can act in adjacent districts without penalty.
  \item \textbf{Prestige} (old name, art) --- \emph{Petition} here \textbf{DV --1} with patricians; \emph{Blackmail} \textbf{+1 SB} against you if scandal hits.
  \item \textbf{Ward-Woven} (sigils, bells) --- \emph{Infiltrate} here starts \textbf{Desperate} unless key is held.
  \item \textbf{Unionized} --- \emph{Strike} is a valid Event; \emph{Broker} with unions here starts \textbf{Controlled}.
  \item \textbf{Shadowed} (underworld ties) --- \emph{Smuggle} \textbf{DV --1}; \emph{Expose} \textbf{Position --1} against you.
  \item \textbf{Water-Right} --- acts as a \textbf{String} for bridges/canals toll negotiation.
  \item \textbf{Archive} --- \emph{Research/Expose} \textbf{DV --1} with proof assembled here.
\end{itemize}

\subsection{Integrity \& Resources}
\begin{itemize}
  \item \textbf{Integrity [6]} ticks from sabotage, audits, disasters, or neglect. At fill: choose \textbf{Shutdown}, \textbf{Seizure}, or a \textbf{Catastrophe} scene.
  \item \textbf{Resources [6]} drop from upkeep, shocks, strikes; refill via Orders, scenes, or Projects. If Resources = 0, future Upkeep requires \textbf{Favor} or a concession clock.
\end{itemize}

\section{Asset Play}
\subsection{Acting Through an Asset (On-Scene)}
Stage a scene at or with an asset; apply its Tags and Strings to Position/DV like any venue. A named staffer may act as a \emph{Follower} or as a cohort (tick \textbf{Integrity} instead of Harm on failures).

\subsection{Upkeep \& Yield (Each Cycle)}
\begin{itemize}
  \item \textbf{Upkeep:} Pay 1--2 of coin/favor/rite. If skipped: tick \textbf{Resources --1} and start \textbf{Creditor [4]} or \textbf{Inspection [4]}.
  \item \textbf{Yield:} If \textbf{Resources $\geq$ 3} and no active \textbf{Accident/Inspection}, gain one: coin, \textbf{Favor (narrow)}, \textbf{Clue}, or \textbf{Leverage (1)} themed to the asset.
\end{itemize}

\subsection{Asset Orders (Between Sessions)}
Choose up to \textbf{2 orders} per session per asset (1 if Resources $\leq$ 2):
\begin{itemize}
  \item \textbf{Operate:} generate Yield with risk (on 1s, tick \textbf{Heat +1} or \textbf{Integrity +1}).
  \item \textbf{Improve:} advance a \textbf{Project} [4--8].
  \item \textbf{Secure:} reduce \textbf{Heat --1} or add a \textbf{Security} tag for one scene.
  \item \textbf{Expand:} start a new Project to add a String (route, permit, office).
  \item \textbf{Audit:} convert one unspent Favor at this venue into \textbf{Standing (Tag)} after proof.
  \item \textbf{Exploit:} trade 1 \textbf{Integrity} for immediate coin + Favor; mark \textbf{Exposure +1} to someone.
\end{itemize}
\paragraph{Resolution} The directing PC rolls once per order (appropriate action). Tier/Tags adjust Position/DV. On hits, apply the order effect; on 1s, GM banks SB and hits Heat/Integrity/Exposure.

\section{Campaign Continuity}
\subsection{Asset Legacy System}
Assets evolve and leave marks across campaigns through their \textbf{Legacy Notes}:
\begin{itemize}
  \item \textbf{Reputation Effects:} Notorious (feared), Respected (trusted), Infamous (recognized), Legendary (historical significance)
  \item \textbf{Historical Events:} Siege Survived, Scandal Weathered, Expansion Completed, Crisis Averted
  \item \textbf{Relationship Changes:} Patron Gained/Lost, Rival Created/Resolved, Community Bond/Feud
  \item \textbf{Physical Evolution:} Renovated, Damaged, Expanded, Relocated
\end{itemize}

\textbf{Legacy Benefits:} Assets with positive legacy gain +1 Resources or +1 Integrity at start of new campaigns.\\
\textbf{Legacy Burdens:} Assets with negative legacy start with Heat +2 or Sanction +1 with relevant patrons.

\subsection{Patron Relationship Continuity}
Worldly Patron relationships persist with modified standing:
\begin{itemize}
  \item \textbf{Active Patrons:} Reduce Tier by 1 but retain 1 Standing tag
  \item \textbf{Dormant Patrons:} Convert Tier to Favor Ledger entries; can be reactivated
  \item \textbf{Former Patrons:} Become Rivals [4] or Contacts [4] based on ending relationship
  \item \textbf{Legacy Patrons:} Historical figures whose mandates still influence current politics
\end{itemize}

\subsection{Campaign Transition Events}
At campaign end/beginning, roll for transition events:
\begin{itemize}
  \item \textbf{Economic Shift (d6):} 1-2 Resources +1, 3-4 No change, 5-6 Resources -1
  \item \textbf{Political Change (d6):} 1-2 Gain new patron opportunity, 3-4 Status quo, 5-6 Lose a patron string
  \item \textbf{Physical Event (d6):} 1-2 Asset improves, 3-4 No change, 5-6 Asset takes Integrity -1
\end{itemize}

\section{Worldly Patrons}
\subsection{Patron Sheet (Template)}
\begin{framed}
\noindent\textbf{[PATRON NAME]}\\
\textbf{Type:} noble / guild / office / temple / factor / satrap / councilor\\
\textbf{Stance toward PCs:} Allied / Wary / Hostile\\
\textbf{Strings (3):} writs, permits, routes, rites, audiences\\
\textbf{Boons:} what they can grant (seals, escorts, stipends, protection)\\
\textbf{Obligations:} tithes, appearances, tasks, ideological lines\\
\textbf{Tracks:}\\
\hspace{10pt}- Favor Ledger (narrow favors owed or granted)\\
\hspace{10pt}- Sanction [4]: warning $\rightarrow$ censure $\rightarrow$ seizure/revocation $\rightarrow$ hunt/prosecution\\
\hspace{10pt}- Patron Exposure [6]: public risk to them from association with you\\
\textbf{Patron Tier:} 0 Contact / 1 Sponsor / 2 Patron / 3 Protector\\
\textbf{Mandate/Crisis Effects:} how their public wins/losses alter your Position/Exposure in their venues\\
\textbf{Legacy Status:} Historical role, ongoing influence, campaign connections
\end{framed}

\subsection{Patron Tiers \& Benefits}
\begin{itemize}
  \item \textbf{Tier 0 --- Contact:} 1 \textbf{Audience}/session; one minor seal once.
  \item \textbf{Tier 1 --- Sponsor:} \textbf{+1 Position} once/scene in their venues; 1 \textbf{Stipend}/session (coin or permit access).
  \item \textbf{Tier 2 --- Patron:} cancel \textbf{one Social SB} per session in public; \textbf{DV --1} on \emph{Petition} to their offices; claim \textbf{Escort} once.
  \item \textbf{Tier 3 --- Protector:} start public scenes \textbf{Controlled} in their venues; \textbf{Endorsement} creates a 4-clock \emph{Bandwagon} on targets.
\end{itemize}
\paragraph{Advance a Tier} Hold \textbf{Standing} with them or convert \textbf{3 favors} into Standing and complete a \textbf{Patron Task [4--6]} on-screen.

\subsection{Obligations \& Sanctions}
\begin{itemize}
  \item \textbf{Obligations:} tithe; keep scandal quiet; show at rites; take contracts; avoid rivals; uphold a doctrine. Skipping two cycles: tick \textbf{Sanction +1}.
  \item \textbf{Sanction [4]:} \emph{Admonish} (lose once/scene Position boost) $\to$ \emph{Censure} (no stipends; add \textbf{Audience: Skeptical}) $\to$ \emph{Seizure/Revocation} (lose a String or asset tag) $\to$ \emph{Hunt/Prosecution} (start \textbf{Warrant [6]}).
  \item \textbf{Patron Exposure [6]:} Your public failures can tick this; at fill they \emph{distance} (Tier --1) or \emph{flip} to Hostile.
\end{itemize}

\subsection{Bargaining Procedure (Negotiation Scene)}
\textbf{Frame:} what boon you want and what obligations you accept.\\
\textbf{Set:} venue tags; your Reputation/Standing apply; use \emph{Petition/Broker/Expose}.\\
\textbf{On hit:} gain the boon; write the Obligation and tick Favor Ledger $\pm$ as appropriate.\\
\textbf{On 1s:} GM may add a secret clause, leak leverage, or start \textbf{Rival Patron [4]}.

\subsection{Multiple Patrons}
You may keep two active patrons without penalty. A third creates \textbf{Split Loyalty [4]}; on fill, one patron issues an ultimatum: choose, public denunciation, or give up a String.

\section{Events, Audits, \& Market Shocks}
Roll or draw 1--2 between sessions per active asset/patronage.
\begin{itemize}
  \item \textbf{Inspection:} start \textbf{Inspection [4]}; on fill, tick \textbf{Integrity +1} or \textbf{Sanction +1}.
  \item \textbf{Accident:} \textbf{Accident [4]} threatens staff; on fill, Integrity +1 and \textbf{Audience: Fearful}.
  \item \textbf{Shortages:} \textbf{Resources --1} and start \textbf{Short Rations [4]}.
  \item \textbf{Tax/Lease Hike:} choose coin cost or \textbf{Sanction +1}.
  \item \textbf{Rival Claim:} assert prior right; start \textbf{Litigation [6]} or \textbf{Duel of Proof [4]}.
  \item \textbf{Rumor Run:} if \emph{Crowd-Facing}, add \textbf{Skeptical}; if \emph{Prestige}, add \textbf{Fascinated} (cuts both ways).
  \item \textbf{Favor Called:} a patron demands service; refuse $\to$ \textbf{Sanction +1}.
\end{itemize}

\section{Blueprints (Projects \& Upgrades)}
Pick a \textbf{Project [4--8]} to install a blueprint; on fill, add the effect and tag.
\begin{itemize}
  \item \textbf{Safehouse (4):} add \emph{Fortified} + \emph{Secret}; gain \textbf{Leverage (1)} once/session from stashed goods.
  \item \textbf{Workshop (6):} add \emph{Hazardous}; once/session, turn coin $\to$ \textbf{Clue/Prototype} tag.
  \item \textbf{Archive Annex (6):} add \emph{Archive}; \emph{Research/Expose} DV --1 here; bank \textbf{1 Clue} each cycle if Upkeep is paid.
  \item \textbf{Shrine-Nave (6):} add \emph{Sanctified}; once/session \emph{Host Rite} starts \textbf{Dominant}.
  \item \textbf{Barge Fleet (8):} add \emph{Mobile} + \emph{Water-Right}; once/session move a scene to adjacent district at same Position.
  \item \textbf{Guard Contract (6):} add \emph{Licensed}; once/session \emph{Call the Watch} auto-succeeds at minor level.
  \item \textbf{Union Hall (6):} add \emph{Unionized}; once/session flip a \textbf{Strike} counter one step toward peace (if dues paid).
  \item \textbf{Playhouse Front (4):} add \emph{Crowd-Facing}; create Audience tags more easily; once/session convert \textbf{Audience: Warm} to \textbf{Favor (narrow)}.
\end{itemize}

\section{Regional Kits (Examples)}
\subsection*{Mid Ahkaz --- Violet Steppe/Meadows}
\textbf{Assets:} Caravan Yard (Mobile, Licensed); Dye Vault (Hazardous, Secret).\\
\textbf{Patrons:} Coin-Weigh Tribunal (Sponsor$\to$Patron), Oasis Clans (Protector with water rites).\\
\textbf{Events:} forged Water Share deeds; desert guides strike; curfew at Steppe Gate.

\subsection*{Ecktoria --- Marble \& Fire}
\textbf{Assets:} Aqueduct Valve Lease (Licensed, Ward-Woven); Foundry Line (Hazardous, Unionized).\\
\textbf{Patrons:} Imperial Exarchate (audit stipends), Legions Remnant (escort writs).\\
\textbf{Events:} water theft panic; audit sweep; relic procession crowds disrupt supply.

\subsection*{Silkstrand --- City of Bridges}
\textbf{Assets:} Tollhouse on Archivolt (Licensed, Prestige); Playhouse (Crowd-Facing, Secret).\\
\textbf{Patrons:} Bridge-Lords (tolls), Playhouse Guild (licenses).\\
\textbf{Events:} fog bell failure; satire lawsuit; Night Keys leak.

\subsection*{Thepyrgos --- Synod \& Collegium}
\textbf{Assets:} Harbor Pilot Office (Water-Right, Licensed); Scriptorium Desk (Archive, Prestige).\\
\textbf{Patrons:} Archons' Synod (writs), Collegium (oath registry).\\
\textbf{Events:} censure threat; procession scandal; tithe curse.

\subsection*{Zakov --- Iron River, Ash Lanterns}
\textbf{Assets:} Ash-Lantern Line (Licensed, Hazardous); Smugglers' Cut Lease (Shadowed, Mobile).\\
\textbf{Patrons:} Ironmasters' Collegium (ore contracts), Lantern Wardens (curfews).\\
\textbf{Events:} ash curfew; lantern line failure; picket riots.

\section{Example of Play (short)}
\textbf{Setup:} PCs lease a \emph{Tollhouse} (Licensed, Prestige) in Silkstrand. Integrity 4/6, Resources 3/6. Patron: \emph{Bridge-Lords} (Tier 1 Sponsor; Obligations: dues, public decorum).\\
\textbf{Scene --- Petition:} They seek a \emph{Tariff Exemption} for a festival barge. Position \textbf{Controlled} (Prestige + patron venue). DV 3. Strong hit: gain the boon; Favor Ledger +1; GM banks SB from a rolled 1 $\to$ \textbf{Rumor Run} starts.\\
\textbf{Orders:} \emph{Operate} (yield coin; a 1 ticks Heat +1). \emph{Improve} (Archive Annex [6] +1 segment).\\
\textbf{Event:} \textbf{Inspection [4]} begins. PCs \emph{Secure} next cycle; on a hit reduce Heat --1. A satirical Playhouse offers help for a cut---risking \textbf{Patron Exposure} if mocked.\\
\textbf{Dilemma:} A rival \textbf{Patron} (Playhouse Guild) offers Tier 2 if they host a scandalous masque; \textbf{Split Loyalty [4]} begins. On fill, Bridge-Lords demand: cancel the masque or surrender \emph{Night Keys (String)}.\\
\textbf{Campaign Transition:} At arc's end, roll Economic Shift (d6: 3 = No change), Political Change (d6: 1 = Gain new patron opportunity), Physical Event (d6: 6 = Asset takes Integrity -1). Tollhouse becomes "Notorious" legacy with "Scandal Weathered" event.

\section*{GM Reference (one page)}
\begin{itemize}
  \item \textbf{Asset Tracks:} Integrity [6] (fill = shutdown/seizure), Resources [6] (0 = shortages), Heat [6] (fill = rough visit/audit).
  \item \textbf{Orders:} Operate \textbullet{} Improve \textbullet{} Secure \textbullet{} Expand \textbullet{} Audit \textbullet{} Exploit.
  \item \textbf{Upkeep/Yield:} pay costs $\to$ choose coin/Favor/Clue/Leverage if safe.
  \item \textbf{Patron Tiers:} 0 Contact \textbullet{} 1 Sponsor (+Pos once/scene; 1 stipend) \textbullet{} 2 Patron (cancel 1 Social SB; DV --1 to Petition) \textbullet{} 3 Protector (start Controlled; Bandwagon clock).
  \item \textbf{Sanctions:} Admonish $\to$ Censure $\to$ Seizure/Revocation $\to$ Hunt/Prosecution.
  \item \textbf{Events:} Inspection \textbullet{} Accident \textbullet{} Shortages \textbullet{} Tax/Lease Hike \textbullet{} Rival Claim \textbullet{} Rumor Run \textbullet{} Favor Called.
  \item \textbf{Continuity:} Legacy effects modify starting conditions; patrons retain Standing; transition events reshape holdings.
\end{itemize}

\vfill
\section*{Changelog}
\textbf{v0.2} --- Added Campaign Continuity section with Legacy System, Patron Relationship Continuity, and Transition Events. Enhanced templates with Legacy Notes and Status fields.\\
\textbf{v0.1} --- First pass with Asset/Patron sheets, tags, orders \& upkeep, patron tiers/obligations/sanctions, events \& blueprints, regional examples, and a worked example of play.

\begin{center}\small This module adds procedures only; defer to the core SRD for roll math and adjudication.\end{center}

\end{document}
