\documentclass[11pt]{book}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a5paper, margin=1.2in}
\usepackage{ebgaramond}
\usepackage{titlesec}
\usepackage{epigraph}
\usepackage{xcolor}
\usepackage{setspace}
\onehalfspacing
\setlength{\epigraphwidth}{0.7\textwidth}
\definecolor{ashgray}{gray}{0.3}
\newcommand{\ledger}[1]{\textsc{#1}}
\newcommand{\ritetag}[1]{\textsf{[#1]}}
\newcommand{\clock}[1]{\ledger{#1 Clock}}

\titleformat{\chapter}{\normalfont\huge\itshape}{\thechapter}{20pt}{}
\titleformat{\section}{\normalfont\Large\scshape}{\thesection}{15pt}{}

\begin{document}

\frontmatter
\title{\Huge\itshape Bitter Root\\\large Tales of Unseen Labor, and the Cords That Bind}
\author{Dusana of the Raven Road}
\date{Under the lamp, bread and salt------the ink is ash and well-water that remembers.}
\maketitle

\chapter*{Preface: The Root That Feeds, the Root That Binds}

\epigraph{The women who wash the dead do not sing. They count. One for the breath that left, two for the name that stays, three for the grief they will carry home. No one thanks them. The soup is free. The price is never asked.}{------Old Mila, hedge- keeper of the Moss Road}

A ledger of stories is a ledger of unpaid hours. The washing, the nursing, the midwifery, the stirring of pots that never empty------none of it appears in the sagas. Heroes do not HEAL fences. Queens do not change linens. And yet the world turns because someone, somewhere, is doing the work that leaves no mark.

This book is not for heroes. It is for the women who stay in the kitchen while the baronet feasts, who walk into the fog to find a lost child, who pour the eighth cup and leave the ninth empty because they know the Hollow drinks last. It is for the rage that does not scream------the rage that braids itself into patience and calls itself \textit{care}.

The first story is a flight. A village burns, a girl runs, and a coven of hedge-witches, Daughter renegades, and a Tulkani storyteller smuggle her across three borders. The second story is a siege by other means: a Hearth Witch and a Daughter of the Chain, locked in a cold war over a baronet's household, who slowly recognize each other's teeth and decide------grudgingly, frostily------to work together.

I have written them in the old way: by a borrowed lamp, on paper that had been a bandage. The mechanics are here, but they are not the point. The point is the weight of the work. The point is the thread.

--- Dusana of the Raven Road, at a crossroads where the well still sings

\mainmatter

\chapter{The Smell of Thyme}
\epigraph{They burned the orchard first. Thyme grows back. Girls don't.}{------Quinn, age twelve, before she forgot her name}

\section*{I. The Root Cellar}

The smell of thyme is the last thing Quinn remembers clearly. Her mother had planted it along the eastern wall, where the morning sun hit the stones just so. \textit{Thyme for courage}, her mother used to say, pinching a leaf between her fingers. \textit{Thyme for memory.}

Then the torches came over the hill.

Quinn is twelve, small for her age, with a mage's knot behind her left ear that she has never learned to name. She can hear the last words of dying plants. It is not a useful gift------until it is. When the wheat screams, you know the fire is close.

The Chain-Lanterns of Thepyrgos move with the precision of men who have done this before. Five of them, hooded in iron mesh, their leader carrying a writ stamped with a patriarchal seal. They do not knock. They kick in the door of the house where Quinn's mother is still cutting bread.

\textit{Where is the girl?}

Her mother says nothing. She is good at nothing. It is the first gift a village witch gives her daughter: the ability to become furniture.

Quinn is already in the root cellar, the door concealed behind a rack of drying herbs. She counts her breaths------eight, never nine, because nine is for the Hollow. Through the cracks in the floorboards, she sees the glint of a Lantern's bell. They ring it once before accusation, three times before burning. She has heard that bell in her nightmares for years. Now it is real.

\section*{II. Old Mira's Last Lesson}

The village witch, Old Mira, is not surprised. She has been expecting the Chain-Lanterns since the autumn, when a neighbor's son saw Quinn whispering to a frost-killed apple tree and heard the tree whisper back. Mira has already sent word to the Breath-less network------a cracked clay bowl left on the north doorstep, a red thread tied around the well rope.

She finds Quinn in the root cellar as the first thatch catches fire.

\epigraph{Draw a line. Any spirit will respect a line if you draw it with witness.}{------Old Mira, her last words}

Mira gives Quinn three things: a piece of chalk, a knot of red thread, and a name that is not her own. \textit{You are not Quinn anymore. You are Thyme. And you will run.}

The Chain-Lanterns' procedure is relentless. They establish a \ledger{Lantern Line} around the village------women's rites cannot cross without test. They demand that every household present its patriarch to swear loyalty. Old Mira, who has no patriarch, who has been the village's memory for forty years, is dragged into the square.

They do not burn her immediately. They question her first, because the Inquisition loves procedure. \textit{Where is the girl? Name her true name.}

Mira looks at the smoke rising from the orchard. The thyme is burning. She smiles.

\textit{She is already gone. And you will not remember her face.}

It is not a spell. It is a curse, and she has been weaving it for years. Every bowl of soup she gave the village, every whispered comfort at a deathbed, every child she delivered------all of it was thread. Now she pulls.

The Chain-Lanterns' leader blinks. He knows Quinn's mother's face, but the daughter's? It slips from his memory like water from a cracked bowl. He writes \textit{girl, approximately twelve, dark hair} in his ledger, and even as he writes, he knows it is not enough.

\section*{III. The Covens' Road}

Quinn------she will learn to answer to Thyme, but not yet------runs through the burning orchard. The thyme smoke is thick and green. It covers her scent. Behind her, the village screams. Ahead, the dark.

The first safe house is a charcoal-burner's hut three miles into the woods. The woman there is not a witch. She is a widow who lost her daughter to a fever and now owes the Breath-less a debt she will never fully repay. She gives Quinn bread and salt, and a new pair of boots that are too large.

\textit{The Hearth-Menders will meet you at the ford.}

The Hearth-Menders are not a coven in the formal sense. They have no patron, no thiasos, no written code. They have clay bowls and lullabies sung in harmonics, and they understand one thing: \textbf{injury obligates repair}. A village that burns owes the survivors a future.

\subsection*{Mechanics: Hearth-Mender Hospitality}
When the Hearth-Menders shelter a fugitive, they create a \clock{Task-Bond} of 4 segments. The fugitive must perform a small act of care------mending a net, telling a true story, watching a child for an hour------before receiving the next safe passage. This is not cruelty; it is the law of mutual aid. \textit{Share your hunger; name your fear; accept a task before a gift.}

Quinn does not understand why she must braid the widow's hair before the widow will give her a cloak. But she does it. Her hands shake. The widow's hair smells of woodsmoke. When she finishes, the widow says, \textit{Now you are not just a girl running. You are someone who gave something. That is harder to forget.}

\section*{IV. The Hearth-Mender's Kitchen}

The next safe house is not a hut but a longhouse deep in the eastern woods, where a circle of Hearth-Menders gathers once a week to grind grain, salt pork, and count the children who have gone missing. Quinn is given a stool by the fire and a sack of beans to sort.

She does not speak their language------not the words, but the rhythm of the work. They do not ask her story. They ask her to stir the pot.

For three days, she sorts beans and listens. The women talk in fragments: a lantern line broken near the ford, a girl smuggled out of Galanina, a Daughter who refused to cut a dying man's thread. They do not whisper. They speak at normal volume, because the Hearth-Menders believe that secrecy is a form of debt------the more you hide, the more you owe.

On the third night, the oldest of them, a woman called Mother Hemlock, sets a bowl of thyme soup before Quinn.

\textit{You can hear plants,} Mother Hemlock says. It is not a question.

Quinn nods.

\textit{That is not a curse. It is a witness. The wheat screamed because it was afraid. You heard it. Now you must tell someone. That is the work.}

She does not explain what Quinn should tell, or to whom. But Quinn understands, in the way that twelve-year-olds understand things they cannot yet name, that her gift is not for hiding. It is for \textit{carrying}.

That night, she whispers the wheat's scream into the hearth. The fire crackles. The women nod. They have heard it before.

\subsection*{Mechanics: The Witness of Plants}
A witch who can hear the last words of dying vegetation may spend a scene in a burned or poisoned field to ask one question: \textit{What killed this place?} The answer is not a name but a direction------a source of blight, a patrol route, a landlord's signature. The GM must answer truthfully, though the truth may be a riddle.

\section*{V. The Daughter's Ferry}

The black-water river is the border. On the far side, the Chain-Lanterns' writ means nothing. But the river itself is a threshold, and thresholds demand tolls.

The ferry is a flat-bottomed boat, its pilot a woman in grey robes with a brass hook for a hand and a child standing beside her------a Hollowed, its eyes like spilled milk. A Daughter of the Cord. A renegade.

Quinn hesitates. Everyone knows the Daughters. The soup is free. The price never is.

\epigraph{Give me a memory. Not a pretty one. The one you would least like to lose.}{------Sister Bryn of the Unbroken Cord (renegade)}

The Daughter's name is Bryn------Sister Bryn, though she no longer wears the cord of the coven. She left the High Mother's service after she was ordered to harvest a dying woman's last breath and the woman had been kind to her. She says the woman died giving her lord husband the son that he demanded. She gave him a daughter two springs before. There is a clock ticking in Bryn's flesh, a \ledger{Redemption} clock that she does not know how to fill.

She rows in silence. The Hollowed watches Quinn with its milk-white eyes. The fog thickens and the sound of oars hitting water fades.

\textit{The smell of thyme,} Quinn says. \textit{That's the memory. My mother's garden.}

Bryn nods. She does not write it down. She does not need to. The memory is now a knot in her own Cord, a debt she will carry until she finds a way to pay it forward.

On the far bank, Quinn stumbles ashore. Bryn does not say goodbye. She turns the boat and rows back into the fog, the Hollowed's hand on her sleeve.

\subsection*{Mechanics: Renegade Daughter's Cord}
A Daughter who breaks from the Cord keeps her Hollowed but no longer owes the High Mother. Instead, she tracks a \clock{Redemption} of 6 segments. Each time she gives shelter or aid without demanding a price, she marks a segment. When the clock fills, she may choose to destroy her Hollowed, rejoin the coven, or transform her Cord into a new kind of bond------one built on mercy rather than debt.

\section*{VI. The Silk-Weaver's Attic}

The last safe house is a Lethai-ar silk-weaver's cottage in the Crimson Valley. The weaver, a woman called \textit{Mask-Quill}, does not speak to Quinn for three days. She simply sets a bowl of soup on the table each morning and takes the empty bowl away each evening.

On the fourth day, she asks: \textit{What is your name?}

\textit{Quinn.}

\textit{No. That name is ash. Choose another.}

Quinn looks at the bowl. The soup is thick with thyme.

\textit{Thyme,} she says. \textit{My name is Thyme.}

Mask-Quill nods. She takes a piece of red silk and ties it around Thyme's wrist. \textit{This is a witness cord. It will remember the truth of you, even when you forget.}

She does not explain what the truth is. Thyme does not ask. She is twelve. She has watched her village burn. She has crossed a black river on a Daughter's boat. She has learned that the world runs on debts she did not incur and must repay anyway.

She will not forget the smell of thyme.

\section*{VII. The Bitter Root}

\epigraph{The root that feeds is the root that binds. The hand that rocks the cradle is the hand that sharpens the knife.}{------from the \textit{Book of the Unnamed}, forbidden in Thepyrgos}

Quinn------Thyme will grow up in the Crimson Valley. She will learn to weave silk and to tie witness cords. She will never see her mother again. She will never know that Old Mira's last curse worked: the Chain-Lanterns' report described a girl with brown hair and no name, and the Inquisition closed the case six months later.

But she will keep the red thread on her wrist. And when she is old enough, she will join the Hearth-Menders. Not out of revenge. Out of \textit{care}. Because someone has to stir the pot.

The soup is free. The second bowl will cost you a story.

\subsection*{Mechanics: Bitter Root Condition}
A character who has survived state violence and been smuggled to safety may acquire the \ledger{Bitter Root} condition. It grants +1 die to acts of mutual aid (feeding, sheltering, healing others) but --1 die to acts of self-advocacy (asking for pay, demanding recognition, refusing a burden). The condition can be \textit{pruned} by an act of witnessed rage------a public accusation, a broken tool, a screamed truth. Pruning removes the penalty and converts the bonus to a permanent \ledger{Coven Bond} String.
\newpage

\section*{The Milestone Mending}
\addcontentsline{section}{section}{The Milestone Mending}
\markboth{THE MILESTONE MENDING}{THE MILESTONE MENDING}

\subsection*{Prologue – The Offering Falters}
The scent of woodsmoke and honeyed mead clung to the air like a promise as Vey stirred the cauldron over the festival fire. Around her, the village square buzzed—looms clacked, children chased darting geese, and the newly milled barley sent up a sweet, dusty haze. But Vey’s eyes kept returning to the standing stone at the square’s heart, where villagers had spent the morning binding its cracks with fresh clay and pressing millet bread into the seams—a \textit{Bread to the Milestone} offering to quiet the Peat‑Fog’s whisper. The stone stood whole again, its three arms (\textit{Regret}, \textit{Hunger}, \textit{The Same}) no longer pointing inward in accusation but outward, like fingers tracing the horizon. A single green shoot pushed through the clay near its base—a sign, she thought, that the Curse was listening, if not yet appeased.

She dipped a finger into the thick, spiced broth lifting from the cauldron—barley, dried apples, a pinch of salt from the Blackwood Charcoalers’ hollow—and touched it to her tongue. The taste was earth, was patience, was the weight of thirty winters spent stitching this village together from the frayed edges of Acasia’s wars. \textit{This} was her true tally‑rod now: not the notched ash‑wood she carried in her youth to mark tithe defaults, but the living count of souls who had chosen to stay, to speak, to mend. \textit{She knew the weight of a broken promise because she had once held the rod that measured them.}

The fire crackled, sending a shower of sparks into the dusk. Vey wiped her hands on her apron, feeling the rough weave ground her in the present. Then she noticed: the clay around the stone’s base was dry and cracking, the millet bread lay in brittle shards, and the green shoot trembled, its tip browned as if scorched. The offering had failed. Some name had gone unsaid.

\subsection*{Scene 1 – Bronn’s Net}
Vey’s thoughts turned to Bronn, the former toll‑collector’s son who now mended the village’s fishing nets. Thirty years ago, he’d arrived with blood still crusted under his nails and a feud burning in his chest like banked coals. His father had doubled the tithe on a starving hamlet during the Grain Blight, demanding payment in iron nails instead of grain; when the hamlet refused, his father had torched their barns. Bronn had carried out the order, his hands shaking as he handed the nails to the weeping widow. The Curse had seized him that night—he’d walked the same mile of road seven times, each loop whispering \textit{Regret} in his ear until he collapsed, frothing at the mouth, begging the stones to remember the names he’d failed to speak.

Vey had found him at dawn, raw‑faced and shivering. She hadn’t offered platitudes. Instead, she’d pressed a lump of cold iron into his palm and said, \textit{“Say their names. Not for the Curse—for them.”} For three days they sat by the river as he named each hamlet family: Elara (yes, he’d named her too, the witch thought wryly), Borin, Tela. With each name spoken aloud, the iron in his hand grew warmer, less like an accursed token and more like a bridge. On the fourth dawn, he walked the road once—and did not return. The Curse had not left him; he still flinched when a bridge creaked underfoot and kept a knotted cord in his pocket to tie when anxiety coiled tight. But today, he laughed as he taught Vey’s granddaughter to splice net, his hands sure and clean, the cord lying forgotten in his toolbox.

Vey approached him where he sat mending a net beneath the awning of the fish‑smoke shed. “Bronn,” she said, keeping her voice low, “the offering failed. Did anything feel off today?”

He looked up, eyes narrowing. “I heard the widow’s boy—Lissa—steal a loaf yesterday. She’s been quiet since. Maybe she’s the one we missed.”

Vey thanked him and moved on, the image of the knotted cord in Bronn’s pocket a reminder that healing left scars, not erasures.

\subsection*{Scene 2 – Davik’s Lute}
Last was Davik, the village’s young bard, whose fingers now danced over the lute with a joy that belied his past. He’d been a courier for a petty lord, sworn to deliver a sealed chest of grain tithes across the Blackwood. Ambushed by feuding lords, he’d fled into the woods, the chest abandoned. The Curse had claimed the road instantly—he’d walked in circles for three days, dehydrated and hallucinating, hearing his lord’s voice accuse him of theft even as he clutched the empty saddle. When he finally staggered out, vow broken and shame burning hotter than fever, he’d tried to hang himself from a mildew‑scarred oak. Vey had cut him down, not with words, but with silence and a shared loaf of bread eaten under the stars.

Her price for his healing had been simple: \textit{“Walk the road with me. Not to deliver—but to see.”} For a month they walked the Imperial Road at dawn, Davik naming every landmark he passed—the burnt orchard, the forked stone, the widow’s elm—until the road stopped feeling like a trap and began to feel like a path. The Curse still touched him: he woke sometimes with the taste of iron on his tongue, and he refused to carry sealed containers, preferring open sacks where he could see the contents. But tonight, he would play for the festival, his lute weaving melodies that spoke of home, not horror. When he glanced at Vey from the stage, his smile held no shadow—only the deep, quiet gratitude of a man who had learned to carry his sorrow without letting it dictate his steps.

Vey found Davik tuning his lute near the firepit. “Davik, did you see Lissa near the stone this morning?”

He shook his head, eyes distant. “I saw her slipping behind the mildew‑wall, clutching something wrapped in cloth. She looked… afraid.”

\subsection*{Scene 3 – The Missing Name}
Vey’s heart hammered. She hurried to the mildew‑wall that bordered the western edge of the square, where the older children sometimes played hide‑and‑seek. There, crouched behind a stack of dried reeds, was a girl no older than twelve, her face streaked with dirt and tears, a small loaf of barley bread clutched to her chest.

“Lissa,” Vey said softly, kneeling to meet the girl’s eyes. “You took the bread without a word. Why?”

The girl’s breath hitched. “My brother… he’s sick. The granary gave us only half a portion today. I… I stole the loaf so he wouldn’t go hungry. I didn’t mean to… I just didn’t know what to say.”

Vey reached out, not to take the bread, but to place her own calloused hand over the girl’s. “You are Lissa of the reed‑wall. You are still here. Say it with me: \textit{Thank you, Lissa.}”

The girl’s voice trembled at first, then grew firmer. “Thank you, Lissa.”

As the words left her lips, the iron warmth Vey felt in her own chest—echoing the iron she had once pressed into Bronn’s palm—spread outward. She glanced toward the standing stone: the clay seemed to drink the evening dew, the millet bread lay whole again, and the green shoot straightened, its tip catching the firelight like a newborn flame.

A flicker of old habit crossed Vey’s mind—the reflex to tally a theft, to mark a debt—but she let it pass, feeling instead the weight of the bread in her hand, the warmth of the stone, and the quiet hum of the festival rising around her.

\subsection*{Epilogue – The Stone Warms}
Vey set the bowl of broth before the moss‑caked base of the stone, where the green shoot now stood tall enough to catch the firelight. Around her, the village celebrated—feet stomped in time to Davik’s lute, children’s laughter rang clear, and the scent of baking bread mingled with woodsmoke. No one spoke of the Curse aloud tonight. They didn’t need to. Its presence was felt in the way Bronn checked a bridge’s knots before crossing, in the way Davik’s eyes flicked to the stone before lifting his lute.

The Curse of Acasia was not cured. It lived in the mildew on old parchment, in the silence between \textit{please} and \textit{thank you}, in the thousand tiny moments where a promise might be broken. But here, in this square of mended stone and shared bread, it was held at bay—not by magic, but by the stubborn, daily choice to see one another, to name the unnamed, to mend what the road had broken. Vey touched the cool stone where the shoot emerged, her heart full. \textit{The ink is wet. The stone is warm. The road will wait—for now.}

\section*{Mechanics: The Milestone Mending}
\addcontentsline{section}{section}{Mechanics: The Milestone Mending}
\markboth{MECHANICS: THE MILESTONE MENDING}{MECHANICS: THE MILESTONE MENDING}

\subsection*{Curse of Acasia: Milestone Offering Failure}
When a \textit{Bread to the Milestone} offering fails (clay cracks, bread shrivels, green shoot withers), the GM secretly notes one unsaid name (a theft, a broken thanks, a forgotten apology). The Curse’s Attention begins to rise.

\begin{description}
\item[Curse Attention Clock] (4 segments) \\
Ticks one segment per failed investigation or per hour of inaction. \\
\textbf{Full}: The standing stone’s arms swing inward (\textit{Regret, Hunger, The Same}); Peat-Fog thickens, whispering forgotten names. Anyone who hears the whisper must make a Resolve roll (DV 3) or begin walking the road in loops (−1 die to all actions until they speak a name aloud). \\
\textbf{Reset}: Speaking the unsaid name aloud with gratitude (e.g., “Thank you, [Name]”) resets the clock to zero and heals the offering (clay seals, bread restores, shoot straightens).
\end{description}

\subsection*{Investigation as Cooperative Threshold Working}
Questioning a villager about the unsaid name is a \textit{Cooperative Threshold Working} (Lore + Survival or Lore + Presence, DV based on villager’s trust).  
\begin{description}
\item[Success] \\
Reduce Curse Attention Clock by 1 segment and gain a clue (e.g., Bronn: “Lissa stole bread”; Davik: “She hid behind the mildew‑wall”). \\
\item[Failure] \\
Gain 1 Hollow Attention tick. The villager becomes withdrawn or defensive (−1 die to future Social rolls with them). Hollow Attention manifests as: \\
\quad • Peat-Fog curling near the questioner’s feet \\
\quad • A sudden chill despite the fire \\
\quad • The sound of a child’s breath just behind the ear
\end{description}

\subsection*{Speaking the Name: Mutual Aid in Action}
When the unsaid name is spoken aloud with genuine gratitude (not coercion), it acts as a Mutual Aid working.  
\begin{description}
\item[Effect] \\
• Resets Curse Attention Clock to zero \\
• The speaker feels iron warm in their chest (+1 die to next Resolve or Empathy roll) \\
• The listener (if the named party) gains the \textit{Seen} condition: +1 die to Social rolls asking for aid until they next break a promise \\
\item[Cost] \\
Nothing is asked—but the speaker must confront their own role in the debt. If they once enforced tithe or measured defaults (like Vey), they take −1 die to the roll (guilt of past enforcement) but gain a permanent \textit{Mender’s String} if they succeed (see \textit{Bitter Root} condition below).
\end{description}

\subsection*{Vey’s Past: Former Enforcer (Condition)}
Vey’s history as a tithe collector shapes her interaction with the Curse.  
\begin{description}
\item[Former Enforcer] \\
• +1 die to Lore rolls involving debt, promises, or broken thresholds \\
• −1 die to Social rolls when demanding payment or enforcing a rule (her old habits clash with her mending) \\
\textbf{Pruning}: If Vey publicly acknowledges her past role in causing harm (e.g., “I once took bread without thanks; I see now it was wrong”), she converts the penalty to a permanent \textit{Mender’s String} (+1 die to all Mutual Aid workings) and removes the −1 die to self‑advocacy.
\end{description}

\subsection*{Household Disposition: Contested Threshold}
The village square during the festival is a \textit{Contested Threshold} (Vey’s mending vs. the Curse’s pull).  
\begin{description}
\item[Effect] \\
• Social rolls investigating the Curse suffer +1 DV \\
• If the party sides fully with Vey’s mending: Dominant (standard DV) \\
• If they try to appease both (e.g., offering bread without naming): Desperate (DV +2) \\
• Disposition shifts only when the Curse Attention Clock is reset or the offering succeeds
\end{description}

\subsection*{Design Notes for GMs}
- \textbf{Clock pacing}: For a 1‑hour festival, tick the Curse Attention Clock every 10–15 minutes of inaction or per failed roll.  
- \textbf{Hollow Attention flavor}: Describe the Peat‑Fog as smelling of wet ash and old parchment; whispers sound like half‑remembered names in the wind.  
- \textbf{Vey’s arc}: If played as an NPC, her Former Enforcer condition creates tension—she’s most effective when she stops measuring and starts mending.  
- \textbf{Player adaptation}: Replace Vey with the players’ healer-type character; give them a similar condition (e.g., \textit{Ex‑Enforcer}, \textit{Debt‑Collector}) to personalize the stakes.  
- \textbf{Failure state}: If the clock fills, the Curse claims one villager (they walk the road looping until rescued via \textit{Search Clock} or \textit{Redemption Clock} workings). The standing stone remains cracked until a full Milestone Offering is made with all names spoken.  
\end{description}

\clearpage
\chapter{The Second Hearth}
\epigraph{She stirred the pot. She watched the pot. They never saw her add the salt.}{------Morwen of Thornwood, on the occasion of no thanks whatsoever}

\section*{I. The New Baronet}

Thornwood is a village of sixty souls, a mill, a ford, and a Hearth Witch named Morwen. For thirty years she has mended bones, settled disputes, and kept the Hollow from the door. She asks nothing but a bowl of soup and a place by the fire. The villagers do not thank her. That is fine. Thanking is not why she does it.

Then the crown appoints a new Knight-Baronet: Sir Aldric of the Eastern Marches, a man with a hunting hawk, a young wife, a sickly infant, and no understanding of hedges. He arrives in a carriage that founders in the ford. His escort tramples the miller's bean field. He does not apologize.

Morwen watches from the kitchen door of her cottage, her thumb pressed against the jamb where the red thread hangs. She already dislikes him.

But it is the wetnurse who makes her blood run cold.

The woman is tall, grey-haired, efficient. She moves like someone who has attended a thousand births and a thousand deaths. She wears no obvious mark of the Chain, but Morwen has been a threshold worker for three decades. She knows the signs: the way the woman never blinks when a baby cries, the way she always pours tea \textit{first for the master}, the way her own set of scales for measuring milk sits beside the hearth like a second altar.

Her name is Sister Bryn.

\epigraph{You're of the Cord. I heard stories.}{------Morwen, the first night}

\epigraph{Stories are debts. I'm not here to pay yours.}{------Sister Bryn, the first night}

\section*{II. The Cold War}

The baronet's household is a small court: Sir Aldric, his wife Lady Margot, newly pregnant, the infant Eleanor, a cook, a groom, and Bryn. Morwen is not part of the household. She is the village witch, which means she is expected to help when things go wrong and to stay out of sight when things go right.

The cold war begins with small things.

Bryn reorganizes Morwen's herb stores \textit{for efficiency}. Morwen misplaces Bryn's witness cords \textit{to air them out}. Bryn salts the well water------just a pinch, not enough to taste, only enough to make Morwen's divinations cloudy. Morwen ties a red thread around the kitchen door------a \ledger{Warm Hand} working that makes the room feel too hot for anyone wearing a Daughter's grey.

Neither woman speaks of it. The baronet notices nothing. His wife, Lady Margot, notices everything but says nothing. She has learned that silence is the only armor that fits.

\subsection*{Mechanics: Household Disposition}
A dwelling under conflicting influences gains a \ledger{Contested} disposition. Threshold workings inside suffer +1 DV. Social rolls involving hospitality are Dominant if you side with one faction, Desperate if you try to please both. The disposition changes only when one witch withdraws or a crisis forces cooperation.

\section*{III. The Broken Fence}

The first crisis arrives not with fire or fever, but with a fence.

The baronet's prized hunting hawk escapes its mews and kills a villager's goat. The villager------a widow named Quinn, weary with an infant granddaughter ------demands compensation. Sir Aldric offers a silver coin and a sneer. Quinn refuses. She goes to Morwen.

\epigraph{Make him pay, witch. He thinks we're nothing.}{------Quinn, her hands still stained with the goat's blood}

Morwen knows she cannot curse the baronet. The Chain-Lanterns are watching------not here, not yet, but they will come if a noble dies of a "witch's fever." She needs another way.

That night, she finds Bryn already standing at the broken fence, running her fingers over the splintered wood.

\textit{You're not going to curse him,} Bryn says. \textit{Too risky.}

\textit{No. But I need him to understand the debt.}

Bryn is silent. Then she takes a length of red thread from her sleeve------Daughter's cord, the kind that records grievances. \textit{A witness cord. He won't see it. His wife will.}

Together, in the dark, they tie the cord around the baronet's carriage axle. Not a curse. A \textit{record}. Every time the axle turns, the cord whispers the widow's name. Sir Aldric will not know why, but he will dream of goats for a week.

In the morning, he sends a purse of silver to Quinn's door. He does not apologize. He does not know why he did it.

Morwen and Bryn do not speak of it. But when Morwen returns to her cottage, the kettle is already warm. Bryn has been there.

\epigraph{You didn't have to help.}{------Morwen}

\epigraph{The goat was not your debt alone.}{------Bryn}

This is the first thread that binds them.

\section*{IV. The Infant's Fever}

Lady Margot's baby, Eleanor, is a sickly thing. She coughs, she does not gain weight, she sleeps too much. Morwen has suspected for weeks that the child has a weak heart------something no poultice can fix, only careful nursing.

One night, the fever spikes. Lady Margot comes pounding on Morwen's door at midnight.

\textit{Please. She's burning up.}

Morwen goes. She brings feverfew, a silver spoon, and her own mother's lullaby. She finds Bryn already there, holding the infant, her face as still as a mask.

The two women work side by side. Morwen cools the child's forehead with a damp cloth while Bryn measures out drops of a grey tincture. Neither speaks. The only sounds are the infant's ragged breath and the soft chime of the baronet's hunting clock in the next room.

At dawn, the fever breaks.

\epigraph{Tincture of willow bark and a drop of your own blood. You're funding the Cord with her survival.}{------Morwen, accusatory}

\epigraph{And your lullaby is a prayer to Inaea. You're binding her to gratitude she'll never be able to repay. Same coin, different face.}{------Bryn, equally accusatory}

They stare at each other across the crib. The infant sleeps, her cheeks pink for the first time in weeks.

\textit{I didn't do it for the debt,} Morwen says.

\textit{Neither did I. But the debt accrues anyway.}

\section*{V. The Missing Child}

The winter that follows is harsh. Margot, her belly large again, is knitting a red blanket. The ford freezes. The mill's waterwheel cracks. Three villagers come down with a lung-sickness that Morwen recognizes as the \ledger{Spreading}------a Daughter's curse, she is sure of it. But Bryn tends the sick alongside her, and the sickness fades without a single death.

Then a village child------a girl named Pella, seven years old, curious and quick------wanders into the woods at twilight and does not return.

The baronet organizes a hunting party. Morwen knows the forest; Bryn can read the Hollow's attention. They go together, leaving the men to thrash about with torches.

\subsection*{Mechanics: Cooperative Threshold Working}
Two witches working toward the same goal can combine their Threshold talents. One rolls \ledger{Lore} (to track the Hollow's signs), the other rolls \ledger{Survival} (to navigate the forest). On a success, they reduce a \clock{Search} clock by 2 segments. On a failure, the Hollow notices them------gain 1 \ledger{Hollow Attention} tick.

They find Pella at midnight, sitting by a standing stone, unharmed. The Hollow was \textit{just looking}, the girl says. \textit{It was lonely.}

Morwen takes the child's left hand. Bryn takes the right. They walk out of the woods without speaking. Behind them, the standing stone hums once, then falls silent.

\section*{VI. The Winter Kitchen}

The baronet throws a feast at midwinter------roasted swan, spiced wine, a fire so large it eats a month's worth of wood. He does not invite Morwen. He does not thank Bryn. He does not notice that the food on the table came from the village's stores, or that the wine was watered, or that the servants are wearing the same coats they wore last year.

In the kitchen, Morwen stirs a pot of soup for the villagers who were not invited. Bryn portions the bread.

\textit{You're not so useless,} Morwen says.

\textit{You're not so sentimental,} Bryn replies.

They drink a cup of small beer. They do not toast. They do not smile. But they do not glare.

\epigraph{The red thread on the kitchen door. Did you tie that?}{------Bryn}

\epigraph{I thought you did.}{------Morwen}

Neither woman unties it.

\section*{VII. The Thaw}

Spring comes. The ford opens. The baronet receives a letter recalling him to court------something about a border dispute, nothing that concerns him. He leaves his wife and child behind. He does not say goodbye to the village.

Lady Margot, to everyone's surprise, asks Morwen to teach her about herbs and knots. \textit{I want to be useful,} she says. Morwen agrees, on the condition that Bryn teaches her about the Hollow in return.

The three women eat together once a week: Morwen, Bryn, and Lady Margot. They do not become friends. Their conversation is the conversation of people who have seen too much to pretend. But they build a small, fragile thing: a community that is not based on blood or marriage or feudal obligation, but on the simple recognition that \textit{someone has to do the work}.

\subsection*{Mechanics: Unseen Labor Clock}
A community that is well tended------by witches, midwives, hedgers, and nurses------maintains a \ledger{Community Health} clock of 6 segments. Each week of care ticks one segment. If the clock is full at the end of winter, the community survives any hardship without loss of life. If it empties, someone starves, or freezes, or sickens beyond help.

Morwen and Bryn, working together, keep the clock full. They never tell the baronet. He never asks.

\section*{VIII. The Final Scene}

One evening, as the last snow melts, Morwen finds a single red thread tied between a chair leg and the pot handle. She knows she did not tie it. She knows Bryn did not tie it.

She leaves it there.

When Bryn comes to fetch the soup pot, she sees the thread. She does not mention it.

But she pours two cups of tea instead of one.

\epigraph{The soup is free,}{------Morwen, pushing a cup across the table}

\epigraph{The price is never,}{------Bryn, taking it}

They drink in silence. The pot simmers. The thread holds.

\chapter*{Colophon}
This book was written on paper that had been a grain sack, in ink cut with ash from a Chain-Lantern's torch. The binding is red thread, tied with eight knots.

The ninth knot is not cut. The work is not done. The women who wash the dead are still counting.

------ Dusana of the Raven Road, at a crossroads where three covens meet

\appendix
\chapter{Mechanics Summary}

\section*{Bitter Root Condition}
\ledger{Bitter Root} (persistent): +1 die to mutual aid actions, --1 die to self-advocacy. Pruned by an act of witnessed rage.

\section*{Unseen Labor Clock}
6 segments. Ticks with weekly care. Full = community survives hardship. Empty = loss.

\section*{Renegade Daughter's Cord}
\clock{Redemption} (6 segments). Ticks when a Daughter gives aid without demanding price. Full grants choice: destroy Hollowed, rejoin Chain, or transform Cord.

\section*{Cooperative Threshold Working}
Two witches combine \ledger{Lore} and \ledger{Survival} to reduce a \clock{Search} clock by 2 on success. On failure, gain \ledger{Hollow Attention}.

\section*{Household Disposition}
\ledger{Contested}: threshold workings +1 DV; social rolls Dominant if siding with one faction, Desperate if trying to please both.
\end{document}
