\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{titlesec}

% Page setup
\geometry{margin=1in}

% Section formatting
\titleformat{\section}{\Large\bfseries\filcenter}{}{0em}{}
\titleformat{\subsection}{\large\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\bfseries}{}{0em}{}

% Item spacing
\setlist{itemsep=0pt,topsep=2pt}
\setlist[enumerate]{label=\Alph*.}

\begin{document}

\begin{center}
{\LARGE\bfseries Black Banners — Condotta \& Crowns}\\[0.5em]
{\large Wargame Expansion for Fate’s Edge}\\[1em]
\rule{\textwidth}{0.4pt}
\end{center}

\section*{I. Introduction: The Theater of Banners}

\subsection*{A. Core Concept}

\paragraph{War as Social Theater.} In Fate’s Edge, battles are brief flashes. The story lives in the hours before and after: payroll arguments, muddy roads, who eats, who prays, who runs. Camps, fords, and counting-tents are the true stages of war.

\paragraph{Mercenaries as Narrative Bridges.} Bannered companies move between courts, ports, and steppes. They tie urban intrigue to caravan legs, wilderness marches, and political endgames. Hire them, join them, outmaneuver them—either way, banners connect arcs.

\paragraph{The Condotta System.} Contracts (condotte) are adventures in legal clothing: clauses, riders, sureties, and scheduled betrayals. This module gives you fast procedures to play the deal, the march, and the clash with the same tools as core Fate’s Edge (Position / DV / Effect, Strings, and clocks).

\subsection*{B. Design Philosophy}

\paragraph{Camp Over Clash.} Put logistics, medicine, and morale on the table. The camp is where choices compound; the clash is where they’re cashed.

\paragraph{Contracts as Stories.} A charter’s margins foreshadow your next session. Every clause is a future scene; every omission, a future wound.

\paragraph{Dynamic Loyalties.} Allegiances shift without spreadsheets. One die per scene nudges fortune, pay, politics, and position.

\paragraph{Cross-Arc Integration.} Banners bring Diamonds (leverage), Clubs (pressure), Hearts (people), and Spades (places) to any arc: court, road, hedge, harbor.

\subsection*{C. Quickstart Summary}

\begin{description}[leftmargin=*,labelwidth=2cm]
\item[Two-Minute Setup]
\begin{enumerate}
    \item Condotta Crossroads: Put two conflicting charters on the table. Ask who you disappoint first.
    \item Company Sheet: Name, Colors, Captain, Cohorts. Add Payday [4], Supply [4], Cohesion [6].
    \item Pressure Clocks: Pick two externals (e.g., Skirmish Spiral [6], Camp Despair [6]).
    \item Allegiances in Flux: At each scene start, roll 1d6 and apply the twist.
    \item Honor ↔ Pragmatism: Slide the dial to reflect recent choices; gain a boon, pay a cost.
\end{enumerate}
\end{description}

\rule{\textwidth}{0.4pt}

\section*{Lore: The Black Banners}

``The contract is the road. The colors are the promise. The pay-chest is the prayer.''

The Black Banners are not one company but a league—a loose commonwealth of mercenary outfits that share camp law, pay customs, and a habit of turning up wherever the map is loud. Their colors are always some device on black: a lantern, a thorn, a bell, a fishbone, a broken scale. Black means coin before crown; witness before oath.

They are the thread that stitches frontiers to courts, caravan roads to siege lines. Ask five people who the Black Banners are and you'll hear five truths: quartermasters who can feed a ghost army; paymasters who count twice; raiders with rules; peacekeepers who charge rent; undertakers who sing.

\rule{\textwidth}{0.4pt}

\subsection*{Origins \& Reputation}

\begin{itemize}
    \item \textbf{Birth at the Crossroads.} The first ``black'' standards reportedly gathered during a famine campaign when city militias and steppe hosts both ran out of grain. A broker chalked clauses on wagon boards; those who signed swore to take coin, keep order, and leave exits. Black cloth was used because no city would lend its colors.
    
    \item \textbf{Four Sacks of Ecktoria.} The empire counts four humiliations as ``sacks.'' The banners will say—privately—that the gates opened for Vilikari cohorts and Banners logistics, not for steppe charges. It was a war of ledgers, ferries, and night roads…and the empire paid late every time.
    
    \item \textbf{Fierce, but priced.} Their reputation is a machine: frightening enough to deter trouble, disciplined enough to be rentable, and bound by camp law that courts grudgingly respect.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Articles of the Camp (the Eight You Hear Aloud)}

\begin{enumerate}
    \item Payday comes before parley. No pay, no move.
    \item Leech and bell are sacred. Healers and chaplains travel untouched.
    \item Colors at dawn, not at night. Sacks and changes of allegiance happen in light, with a witness.
    \item Prisoners eat. Ransom or release; no starving.
    \item The broker is a bridge. Don't kill bridges.
    \item Take no road you won't defend. (Some add: ``Leave two exits; take one.'')
    \item No fire in the hospital quarter. Even if the enemy runs through it.
    \item Count twice. Coin, bodies, promises.
\end{enumerate}

Breaking an Article earns a Black Mark—a stigma that travels faster than mail. Three Black Marks and no reputable broker will carry your charter without triple surety.

\rule{\textwidth}{0.4pt}

\subsection*{How They See War}

\begin{itemize}
    \item \textbf{Camp before clash.} A battle is the interest on debts a captain accrued in camp: food, route, weather, rumor. If those sums are wrong, steel won't save you.
    
    \item \textbf{Contracts as maps.} A condotta (charter) isn't just payment—it's terrain in sentences: rights of passage, shares of salvage, witness duties, who feeds prisoners and at what rate. Good captains can march through a clause.
    
    \item \textbf{Loyalties in motion.} They do change sides, but by rule: daylight, drum, witness, surety paid. The gutter word is ``treachery.'' The banner word is ``rider fulfilled.''
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{The Four Faces of a Banner (Hearts)}

\begin{itemize}
    \item \textbf{Captain.} Judge of risk and reputation. The best can smell a forged clause and a storm in the same breath.
    
    \item \textbf{Quartermaster.} The real warlord. Controls Supply, hates surprises, loves rain barrels.
    
    \item \textbf{Surgeon/Leech.} Holds morale together with needles, vinegar, and songs. Chooses who lives when coin runs out.
    
    \item \textbf{Scout/Sergeant.} Decides if today is a straight road or an honest ambush.
\end{itemize}

If Hearts are rotten, no Spade (place), Club (pressure), or Diamond (writ) will save you.

\rule{\textwidth}{0.4pt}

\subsection*{Colors, Devices, \& Cant}

\paragraph{Common devices:} lantern, bell, thorn, chain, fishbone, ladder, broken bit, river-stone, empty ring.

\paragraph{Cant phrases:}
\begin{itemize}
    \item ``Black's the bargain.'' (We do this for pay, not love.)
    \item ``Lantern high.'' (We act under witness.)
    \item ``Count on copper.'' (Do it the hard way; stop promising silver.)
    \item ``Two weights, one truth.'' (Say the public price and the real one.)
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Famous Companies \& Rumors}

\begin{itemize}
    \item \textbf{The Iron Lanterns.} Siege specialists who promise no fire in hospitals and keep it. They claim to have taken five cities with fewer dead than one bad winter.
    
    \item \textbf{Thornwrights.} Sappers and road-cutters who price bridges by the breath—your breath at the gallop.
    
    \item \textbf{Bone \& Bell.} Undertaker company; they bury both sides by the book and take payment in coin, land, or lineage favors.
    
    \item \textbf{The Fifth Pay-Chest.} An administrative company that ``escorts'' money and occasionally a colonel's conscience.
    
    \item \textbf{Sea Charters.} Colors on black sails; swear by Blue Moot tariffs; infamous for opening blockades with paperwork and weather.
\end{itemize}

\textbf{Rumor:} A captain called the Bannerless One walks with no colors and settles Payday by sheer presence. Those who accept their help wake with coin…and a debt that isn't money.

\rule{\textwidth}{0.4pt}

\subsection*{Brokers, Prefects, and Other Power}

\begin{itemize}
    \item \textbf{Condotta Brokers.} They draft the charters, carry news, hold surety. Killing a broker is like burning a bridge: possible, unforgettable, and expensive for a decade.
    
    \item \textbf{River Prefects.} Stamps that open bridges—once. Prefects sell time more than passage; banners pay dearly.
    
    \item \textbf{City Factors.} Neutral accountants; they read your ledger aloud and write you back into polite society—for a price and a favor. Many banners owe them Price of Neutrality.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Faith \& Rites}

\begin{itemize}
    \item \textbf{Bell-Tokens.} Camp chaplains carry small bells strung with bone or copper. Three bells rung at dawn means a general amnesty for the morning: bury dead, trade prisoners, fix wells, no ambush. Violating a Bell Dawn is how legends end badly.
    
    \item \textbf{The Surgeon's Right.} A leech can claim any blade or cloth in arm's reach during a procedure. Interfere and the nearest captain is obligated to cut you down or lose face.
    
    \item \textbf{Break-Reins Rite.} When leaving a bad charter, a captain hangs a bridle ring on the camp gate. It means: we break this without treachery; pay our arrears, we'll bury your dead. Aveh's riders sometimes attend.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Money, Loot, \& Shares}

\begin{itemize}
    \item \textbf{The Cut.} Standard shares: captain 3; officers 2; soldier 1; specialist +½. Horses get fed first, surgeons second, then soldiers, then officers, then captains. (Good companies keep it that way.)
    
    \item \textbf{Black Chest.} An emergency pay-chest sealed under four keys—captain, quartermaster, leech, and the senior sergeant. Opening it without witness earns a Black Mark.
    
    \item \textbf{Repair Years.} When a sack goes lawful but ugly, some companies pay in repair years—labor, stone, seed—rather than coin. It buys reputations you can recruit with.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Law \& Punishment}

\begin{itemize}
    \item \textbf{Articles tribunals.} Verdicts delivered at noon, with the lantern high. Penalties range from dock of shares to branding (a small black banner on the forearm) to drumming out (colors stripped, tools kept).
    
    \item \textbf{The Gentle Noose.} A derisive term for ``polite'' executions: a leather loop drawn until sleep—then a posting to the worst job for a year. Leechs hate it. Prefects love it.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Relationships \& Politics}

\begin{itemize}
    \item \textbf{With Ykrul.} Respect and rivalry. Banners can pay for route rights with repair or coin; Ykrul will grant them if you hold your exits honest. Many Ykrul learned to price routes from banner quartermasters—and many quartermasters learned geometry from Ykrul judges.
    
    \item \textbf{With Vilikari.} Symbiosis. Vilikari run night ferries, speak to both ledgers, and make false orders less profitable than a right bribe. Every good banner captain owes a Vilikari countkeeper at least one feast.
    
    \item \textbf{With Ecktoria.} The empire hires, audits, and denounces—often in that order. Ecktorian ``corrections'' improve Position but put Black Marks on captains who say no.
    
    \item \textbf{With Aeler.} Hire for stone and siege; pay in repair contracts and courtesy. Break a keystone charter once and all mountain roads become expensive.
    
    \item \textbf{With Lethai.} Tricky. Lethai-al will deal if you respect hedges, bell dawns, and seed dues. Lethai-ar offer night work at good rates and terrible luck.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Seasons \& Omens}

\begin{itemize}
    \item \textbf{Winter:} loyalty crystallizes or shatters. Good captains winter under charter; bad ones invent taxes.
    
    \item \textbf{Storm Season:} the Blue Moot opens sea lanes or closes them. Sea Charters thrive; everyone else writes smaller letters.
    
    \item \textbf{Omens:} a lantern that won't light before parley; a bell that rings once with no wind; a pay-chest key that twists warm in a pocket.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{What Everyone Knows / What Captains Know / What Only Leeches Say}

\begin{description}[leftmargin=*,labelwidth=4cm]
    \item[Everyone:] They fight for pay but keep rules that make them safer than warlords.
    \item[Captains:] The real battle is Payday and Supply; break neither and you can lose three skirmishes and still win the war.
    \item[Leeches:] If the camp sings, you live. If they stop singing, even the best charter bleeds.
\end{description}

\rule{\textwidth}{0.4pt}

\subsection*{Adventure Sparks}

\begin{itemize}
    \item \textbf{Lantern High at Noon.} A broker calls a bell dawn to arbitrate a massacre blamed on your colors. Find the forger, save the Articles, or carry a Black Mark forever.
    
    \item \textbf{The Fifth Chest.} Someone stole the black chest and left everyone's keys in their boots. Recover it before Cohesion unravels.
    
    \item \textbf{Price of Neutrality.} City factors demand a favor: escort a despised official through your own angry camp during Short Rations.
    
    \item \textbf{The Bannerless One.} A quiet person clears your Payday problems and asks only for a prisoner you'd rather keep. Choose mercy, honor, or coin—then count the cost.
    
    \item \textbf{Bridge by Bell.} Blue Moot offers one bell-token to reopen a war-closed bridge for a watch. Who crosses first writes a season's history.
\end{itemize}

\rule{\textwidth}{0.4pt}

\subsection*{Using the Black Banners in Your Campaign}

Drop a bannered company anywhere you need stakes with rules. They carry their own courts, hospitals, road crews, and songs. In intrigue arcs, they're leverage; in caravan legs, they're escorts or competitors; in wilderness play, they keep the wolves fed and the fires small; in political campaigns, they are the deniable arm of policy, standing under black to make the colors behind them look clean.

The secret of the Black Banners is simple: they turn war back into agreements. And agreements—like bridges—are only as strong as the people who keep them in the wind.

\section*{II. Core Mechanics Framework}

\subsection*{A. The Four-Suit System}

\paragraph{Hearts (People).} Faces for your theater:
\begin{itemize}
    \item Captain: command, parley, risk.
    \item Quartermaster: supply, pay, scavenge.
    \item Surgeon/Leech: harm control, disease, morale.
    \item Scout/Sergeant: terrain, ambush, tempo.
\end{itemize}
Use Hearts to bind scenes to stakes: who bleeds if this goes wrong?

\paragraph{Spades (Places).} Position and Effect live in terrain: frozen fords, rain-cut roads, half-built siegeworks, forest lanes under truce. Spades name the rule of the ground and what it does to dice.

\paragraph{Clubs (Pressures).} The SB menu of war: false orders, cut lines, camp-fever, gear breakage, rival aid that arrives late and loud, political “help.” Clubs are how the Keeper spends setbacks.

\paragraph{Diamonds (Leverage).} Fictional assets that bypass a roll once when you pay the price: safe-conducts, bridge priority stamps, debt notes, dispensation writs, blood-truce witnesses. Put at least one Diamond in every scene.

Use in play: At the top of a leg, draft one of each suit. Highest rank sets a major clock (4/6/8/10). Any Ace = winter \& loyalty motif: expect hunger, omen, promotion-by-attrition.

\subsection*{B. Company Sheet Structure}

[COMPANY NAME] — [COLORS/BANNER]\\
Captain: [NAME] \hfill Ledger: What we owe / Who owes us\\
Cohorts: Scouts • Sappers • Surgeons • Foragers • Skirmishers

\textbf{TRACKS:}
\begin{itemize}
    \item Payday [4] — missed pay, bad terms, seized chests → desertion risk
    \item Supply [4] — weather, marches, sabotage → worse Position on maneuvers
    \item Cohesion [6] — losses, betrayals, humiliation → mutters → fractures
\end{itemize}

Ledger Notes (Strings): broker’s favor • bridge priority • chapel bell-token • banner voucher • ford right (season). These are receipts you can spend.

\subsection*{C. Track Management}
\begin{itemize}
    \item \textbf{Payday.} Tick for missed coin, broken promises, stolen chest, or a “creative” delay. On fill: a squad peels off, a veteran sells secrets, or the next Allegiances roll is with disadvantage (take the worse of two).
    \item \textbf{Supply.} Tick for hard weather, forced march, siege diet, or sabotage. While Supply ≥ 3, all risky maneuvers start one step worse Position unless you bring a Diamond to bear (scavenge, requisition, dispensation).
    \item \textbf{Cohesion.} Tick for public humiliation, ugly losses, officer feuds, or moral injury. When Cohesion hits 0–1, the Keeper may convert the first 1 SB each scene into Desertion, Brawl, or Refusal until you repair.
\end{itemize}

\rule{\textwidth}{0.4pt}

\section*{III. Loyalty and Allegiance Systems}

\subsection*{A. Allegiances in Flux (1d6, scene start)}
\begin{enumerate}
    \item Condotta Flip. Rival charter offers better terms. Take it: gain Black Mark [1] with the old patron. Refuse: tick Payday.
    \item Ykrul Counter-offer. Accept an exit-rich but costly route: start Obligation [2] you must pay in concessions later; refuse and tick Payday.
    \item Vilikari Shortcut. +1 Effect on this maneuver via night roads; Keeper gains 1 banked SB to spend as Hostile Terrain this scene.
    \item Ecktorian “Correction.” March under “imperial guidance”: +1 Position this scene or take a Black Mark [1] for defiance.
    \item False Orders. Clubs rain down. First command this scene risks misinterpretation unless countered with a Diamond (cipher, cantor, witness).
    \item Bannerless One. A myth-wrapped fixer clears Payday by 1—but starts Price of Mercy [2]. It will come due (prisoner release, repair years, or a refused sack).
\end{enumerate}

Note: Allegiances events are biases, not dictates. They tilt Position/DV, start clocks, or offer bargains.

\subsection*{B. Honor vs. Pragmatism Dial}

[HONOR] ←— 0 —→ [PRAGMATISM]

\begin{itemize}
    \item Honor (left). +1 Position in parley/oaths; logistics burn faster (tick Supply on a partial where you’d normally be safe).
    \item Pragmatism (right). +1 Effect in ambush/sabotage; social starts one step worse Position when oaths/witnesses matter.
\end{itemize}

Centering Actions: pay arrears on-screen; return prisoners with witness; fulfill a broken writ at cost. Move the dial one step toward center.

\subsection*{C. Loyalty Management}
\begin{itemize}
    \item \textbf{Black Marks.} Track public betrayals and insulted patrons. At 3, your Diamonds cost more: each use adds Political Pressure [1].
    \item \textbf{Favor Debts.} Who owes you? Who do you owe? Record as Strings. Each unpaid Favor at arc’s end ticks Cohesion once (soldiers smell unpaid debts).
    \item \textbf{Reputation Shifts.} Recruiting in civilized ports favors Honor; hiring in hard fronts favors Pragmatism. Slide the dial to attract what you want.
\end{itemize}

\rule{\textwidth}{0.4pt}


\rule{\textwidth}{0.4pt}

\section*{IV. Camp Life Procedures}

\subsection*{A. Weekly Entanglements (choose 2 per leg)}

\begin{enumerate}
    \item \textbf{Condotta Brokers' Syndicate}
    \begin{itemize}
        \item Register colors and surety to bank Broker's Favor (parley DV −1 in camp).
        \item Skip registration: gain Unauthorized Colors—the Keeper can spend the first SB at a checkpoint as False Orders.
    \end{itemize}
    
    \item \textbf{Camp Quartermasters}
    \begin{itemize}
        \item Accept audits: cancel the next Supply tick on this leg.
        \item Refuse: start Short Rations [4]; while active, marching starts one Position worse.
    \end{itemize}
    
    \item \textbf{Field Chaplains \& Leech-Circles}
    \begin{itemize}
        \item Tithe time or coin: clear Camp-Fever [1] each day, and the Keeper cannot convert the first 1 SB into Disease/Death this leg.
        \item Neglect: on the first Clubs spend, camp-fever blooms.
    \end{itemize}
    
    \item \textbf{Steppe Envoys}
    \begin{itemize}
        \item Keep rites (salt/water/shade): bank Remount Rights (once: +1 Effect on retreat/rapid redeploy).
        \item Break rites: gain Soured Oath (parley DV +1 with steppe peoples) until you repair with a gift or witness.
    \end{itemize}
    
    \item \textbf{Hill Truce-Holders}
    \begin{itemize}
        \item Stones-circle witness: bank Blood-Truce (cancel one hill ambush or demand parley).
        \item Violate: begin Guest Right Forfeit [4]; when it fills, locals refuse aid or safe passage.
    \end{itemize}
\end{enumerate}

\subsection*{B. Soft-Power Integration}

\begin{itemize}
    \item \textbf{River Prefect Stamps.} Spend to claim Bridge Priority (crossing DV −1 once).
    \item \textbf{Neutral City Factors.} Reroll a market roll in-camp; start Price of Neutrality [4] (the city expects something back).
    \item \textbf{Border Monasteries.} Bell-tokens make the first Prisoners/Parley step safer (Position +1). Burn a bell-token for a clean retreat once/arc.
\end{itemize}

\rule{\textwidth}{0.4pt}

\section*{V. Combat and Campaign Integration}

\subsection*{A. SB Menu for War (Keeper)}

When you roll 1s or on partials/misses, consider:
\begin{itemize}
    \item False Orders (split unit, wrong target, delay).
    \item Ally Turns Sides (for a scene or for pay).
    \item Cut Supply Lines (tick Supply).
    \item Camp-Fever Outbreak (start or advance disease).
    \item Gear Failure (downgrade Effect; force scavenge).
    \item Reinforcements Arrive (tilt Position against you).
    \item Authority Intervention (political demand interrupts ops).
    \item Winter/Loyalty Omens (hard weather; Allegiances roll with disadvantage).
\end{itemize}

\subsection*{B. Position/DV Modifiers (fast dials)}

\begin{itemize}
    \item \textbf{Weather:} Snow (−1 Position), Storm (DV −1 to cross water), Clear (+1 Effect to ranged).
    \item \textbf{Supply:} Short (−1 Position on risky maneuvers), Adequate (normal), Surplus (+1 Effect on prepared actions).
    \item \textbf{Morale:} Low (DV −1 to resist fear), Normal, High (+1 Position to hold ground).
    \item \textbf{Terrain:} Familiar (+1 Position), Neutral (normal), Hostile (−1 Effect unless you bring a local Diamond).
\end{itemize}

\subsection*{C. Campaign Clock Management}

\paragraph{Primary:}
\begin{itemize}
    \item Payday [4] — financial pressure.
    \item Supply [4] — logistical pressure.
    \item Cohesion [6] — morale pressure.
\end{itemize}

\paragraph{Secondary:}
\begin{itemize}
    \item Skirmish Spiral [6] — tit-for-tat escalates into a war neither side budgeted.
    \item Camp Despair [6] — rumor, grief, and boredom eat discipline.
    \item Political Pressure [4] — courts, patrons, and prefects add new Prices to Diamonds.
\end{itemize}

\rule{\textwidth}{0.4pt}

\section*{VI. Scenario Seeds and Scores}

\subsection*{A. Contract Negotiations — Broker's Table, Two Ledgers}

Two charters: one legal and thin, one dirty and rich. Pick one; secret the other as leverage (String). Attempt a synthesis by proposing a repair project that satisfies both—on a miss, start Political Pressure [2] and tick Payday.

\subsection*{B. Crisis Management}

\paragraph{Plague Ground Sanctification.} Camp-Fever [3/4]; coordinate chaplains, leeches, and a bell-token circle to reset. Each delay ticks Cohesion; success restores one step toward Honor.

\paragraph{Banner Swap in the Wind.} Colors change mid-skirmish; prove command while False Orders fly. Claim a Diamond (cipher, cantor, or witness) to hold line-of-command.

\subsection*{C. Leadership Challenges — Field Promotion}

A shattered unit's warrant lands in your lap. You inherit their Payday [2/4], Supply [1/4], and Black Mark [1]. Keep them intact for a leg and you gain Veteran Network (a permanent contact).

\rule{\textwidth}{0.4pt}

\section*{VII. Cross-System Integration}

\subsection*{A. Political Intrigue}

Map condotta clauses to Influence ticks. Broker's Favor spends like a Boon. Black Marks behave like Exposure in social arcs.

\subsection*{B. Caravan Operations}

Quartermaster audits cancel the next Supply tick on a leg. River stamps reduce crossing DV. Vilikari shortcuts give +1 Effect and bank a hostile-terrain SB for the Keeper.

\subsection*{C. Wilderness Survival}

Hill stones-circles serve as safe trailheads (bank a Blood-Truce). Steppe rites (salt/water/shade) affect travel Position. Environmental hazards tick Supply instead of dealing Harm outright.

\subsection*{D. Urban Adventures}

City factors reroll a market roll; guild dispensations act as Diamonds. Municipal politics can rewrite the fine print on your charter mid-arc.

\rule{\textwidth}{0.4pt}

\section*{VIII. Character Options and Roles}

\subsection*{A. Company Roles}

\paragraph{Captain (leadership).}
\begin{itemize}
    \item Command, parley, and dial control.
    \item Once/scene: convert a No into a Project [4] if you can offer a witness.
\end{itemize}

\paragraph{Quartermaster (logistics).}
\begin{itemize}
    \item Supply, audits, requisitions.
    \item Once/leg: cancel one Supply tick by cashing a String or finding salvage.
\end{itemize}

\paragraph{Scout/Sergeant (tactics).}
\begin{itemize}
    \item Recon, ambush, terrain leverage.
    \item Once/scene: treat hostile ground as neutral if you describe the route.
\end{itemize}

\paragraph{Surgeon/Leech (support).}
\begin{itemize}
    \item Harm, disease, morale rituals.
    \item Once/session: downgrade a fatal consequence to a clock the table can work.
\end{itemize}

\subsection*{B. Specialized Talents}

\begin{itemize}
    \item \textbf{Condotta Veteran (4 XP).} +1 die on contract scenes; spot a forged clause on sight.
    \item \textbf{Banner Reader (3 XP).} Read enemy disposition from standards; predict one Allegiances result per leg.
    \item \textbf{Supply Master (5 XP).} Reduce Supply by 2 segments when you secure forage or plunder; scavenge rolls gain Position +1 in hostile terrain.
\end{itemize}

\rule{\textwidth}{0.4pt}

\section*{IX. Campaign Framework}

\subsection*{A. Session Zero — Why This War?}

``Ecktorian treasuries bankroll a perpetual border-war they claim to `contain,' while companies sell steel to whomever keeps the books. Between rigid honor codes and steppe pragmatism, mercenaries survive by reading winds, not banners.''

\subsection*{B. Arc Progression}

\paragraph{Act I:} Form up, take the first cheap charter, meet the brokers.

\paragraph{Act II:} Run lean, juggle loyalties, pay or bleed for Diamonds.

\paragraph{Act III:} A crisis demands a side; move the dial hard and live with the ledger.

\subsection*{C. Legacy Options}

\begin{itemize}
    \item \textbf{Company Reputation} (affects future rates and riders).
    \item \textbf{Veteran Network} (NPCs become contacts across arcs).
    \item \textbf{War Stories} (Strings you can cash in other modules as legend).
\end{itemize}

\rule{\textwidth}{0.4pt}

\section*{X. GM Toolkit and Procedures}

\subsection*{A. Scene Start Procedure}

\begin{enumerate}
    \item Roll Allegiances in Flux (1d6).
    \item Apply the twist (tilt Position/DV, start a clock, or offer a bargain).
    \item Check Payday/Supply/Cohesion; surface any on-3+ effects.
    \item Set baseline Position from terrain and the Honor/Pragmatism dial.
    \item Put one Diamond on the table (named, with a clear price).
\end{enumerate}

\subsection*{B. Pressure Management}

\begin{description}[leftmargin=*,labelwidth=2cm]
\item[Low:] One primary clock in view; Diamonds plentiful.
\item[Medium:] Two clocks pressure; Clubs spend once/scene.
\item[High:] Three clocks hot; Allegiances roll with disadvantage; require witness to make anything stick.
\end{description}

\subsection*{C. Keeper Cheatsheet}

\begin{itemize}
    \item Allegiances every scene.
    \item Payday/Supply visible at all times.
    \item Offer fiction-first Diamonds.
    \item Spend SB as: False Orders, Ally Turns, Supply Cut, Disease.
\end{itemize}

\subsection*{D. Pacing Guidelines}

A good banner session breathes:
\begin{itemize}
    \item 2–3 camp scenes (contracts, audits, rites).
    \item 1–2 travel scenes (marches, scouts, weather).
    \item 1 major encounter (battle, crisis, betrayal) with at least one Diamond on the table.
\end{itemize}

% ===========================
% Condotta & Crowns — Followers (Cap 1–5)
% ===========================
% ===========================
% Condotta & Crowns — Followers (Cap 1–5) — Refined
% ===========================

\section{Followers: Condotta \& Crowns}
\label{sec:banners-followers-cap}

\subsection*{Cap, Scale, Tracks — Core}
\textbf{Cap (1–5)} measures capability (training, kit, doctrine). \textbf{Scale} is size: I (handful), II (squad), III (wing), IV (mixed-arms company).
\begin{itemize}[leftmargin=*]
  \item \textbf{Cap Edge:} If your unit’s Cap exceeds opposition by 1+, choose \emph{Position +1} \emph{or} \emph{Effect +1}. If lower by 1+, take \emph{Position –1} \emph{or} \emph{Effect –1}. A 2+ edge shifts \emph{both}.
  \item \textbf{Scale Effects (explicit):}
    \begin{itemize}
      \item \emph{Against smaller targets (you \textgreater\ target by 1+ Scale):} take \textbf{Effect +1}; if 2+ Scales larger, also \textbf{DV –1} on area/control orders.
      \item \emph{Against larger threats (you \textless\ target by 1+ Scale):} take \textbf{Effect –1}; if 2+ Scales smaller, also \textbf{DV +1} unless you use precision orders (Recon, Sabotage, Extract).
    \end{itemize}
  \item \textbf{Follower Tracks:} \emph{Loyalty [6], Morale [6], Condition [3]}. Companies also interact with \emph{Payday [4], Supply [4], Cohesion [6]}.
\end{itemize}

\paragraph{Cap Progression (how it rises)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Training Project [6]:} Raise Cap by +1 (max 5) with drill + doctrine + kit refresh; costs 1 \emph{Payday} and 1 \emph{Supply} on completion.
  \item \textbf{Kit Upgrade [4]:} No Cap change, but treat the unit as \emph{Cap +1 for Effect} until the next \emph{Supply} tick.
  \item \textbf{Battle-Hardened (milestone):} After two strong-hit deployments without Condition loss, mark \emph{Veteran}. Veterans resist one SB (see below) per scene.
\end{itemize}

\paragraph{Follower \(\leftrightarrow\) Company Track Interactions}
\begin{itemize}[leftmargin=*]
  \item \textbf{Bleed Up:} If a follower ends a leg at \emph{Condition \(\ge\) 2}, reduce \emph{Cohesion –1} (once per leg).
  \item \textbf{Low Loyalty:} When a follower drops to \emph{Loyalty 1–2}, mark \emph{Cohesion –1} or begin \emph{Mutiny [2]} (Keeper’s choice, fiction-led).
  \item \textbf{Low Morale:} If \emph{Morale 1–2}, their first Order each scene suffers \emph{Position –1} unless \emph{Payday} is cleared or \emph{Chaplain/Leech} time is tithed.
  \item \textbf{Thin Company Tracks:} If the Company hits \emph{Payday [4]} or \emph{Supply [4]}, each deployed follower tests \emph{Loyalty –1} on a partial/miss.
\end{itemize}

\paragraph{Costs (mechanical teeth)}
At \emph{End of Leg}, unpaid \textbf{Costs} (shares, tithes, kit, audits) trigger:
\begin{itemize}[leftmargin=*]
  \item \textbf{Step 1 (warning):} \emph{Loyalty –1}. If already at 1–2, also \emph{Morale –1}.
  \item \textbf{Step 2 (arrears):} Mark a \emph{Black Mark} on the Company (harder hires; polite courts treat Cap –1 for parley).
  \item \textbf{Step 3 (default):} Begin \emph{Desertion [4]} for that unit or lock their Bandwidth (cannot take Orders) until paid.
\end{itemize}

\paragraph{SB Resistance (Veteran/Elite hardening)}
\begin{itemize}[leftmargin=*]
  \item Units with \textbf{Cap 4–5} or \emph{Veteran} may \emph{ignore} one \textbf{False Orders} \emph{or} \textbf{Ally Wavers} SB spend \emph{per scene}. Mark a small \(\diamond\) by the unit to show the resistance is unspent.
\end{itemize}

\paragraph{Quick Reference (Cap/Scale vs Opposition)}
\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Matchup} & \textbf{Adjustment} & \textbf{Example}\\
\midrule
Cap 3 vs Cap 2 & Choose Position +1 or Effect +1 & Vet Scouts vs Bandits \\
Cap 2 vs Cap 4 & Choose Position –1 or Effect –1 & Regulars vs Elites \\
Scale II vs I  & Effect +1 (DV –1 on area) & Platoon vs Team \\
Scale I vs III  & Effect –1 (DV +1 unless precision) & Squad vs Wing \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Order Economy (When to use…)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Hold/Grind:} When you need time/anchor; best with high Scale, high Cap.
  \item \textbf{Screen/Recone:} When information reduces DV later; best early, any Scale.
  \item \textbf{Raid/Sabotage:} When enemy Scale is higher; precision avoids DV +1.
  \item \textbf{Extract/Disengage:} When clocks are hot; trades Effect for clean exits.
  \item \textbf{Seal/Control:} When the battle is won and you want revenue/safe lanes.
\end{itemize}

% ===== TEMPLATE (updated fields) =====
\subsection*{Follower Sheet (Template)}
\begin{center}
\begin{tabularx}{\textwidth}{|X|}
\hline
\textbf{[UNIT NAME]} — Colors/Banner: \_\_\_\_ \\
Type: Bodyguard / Strike Force / Division / Company \quad Playbook: Combat Ally / Specialist \\
\textbf{Cap:} 1–5 \quad \textbf{Scale:} I/II/III/IV \quad \textbf{SB Resist:} \(\square\) (Vet/Elite only) \\
Capabilities: \_\_\_\_ \\
Edges: \_\_\_\_ \quad Liabilities: \_\_\_\_ \\
Strings: \_\_\_\_ \quad Costs: \_\_\_\_ \\
Tracks: Loyalty [\_/6], Morale [\_/6], Condition [\_/3] \\
Tags: \_\_\_\_ \\
Current Deployment: \_\_\_\_ \quad Active Orders: \_\_\_\_ \\
\hline
\end{tabularx}
\end{center}

% ===== SAMPLE UNITS (fields trimmed for brevity) =====
\subsection{Elite Bodyguards}

\begin{followerbox}{Captain's Guard}
Cap: 4 \quad Scale: I \quad SB Resist: \(\square\) \\
Playbook: Combat Ally \\
Capabilities: Close protection, command, last-man extraction \\
Edges: Veteran drills; duelists \quad Liabilities: Oath scruples \\
Strings: Captain’s favor; court dispensation \quad Costs: Shares; oath service \\
Tracks: Loyalty 5/6, Morale 4/6, Condition 0/3 \\
Tags: Elite Training, Oath-Bound, Honor Guard \\
Current Deployment: Staff HQ \quad Active Orders: Interpose
\paragraph{Orders}
\textbf{Interpose} (protect principal); \textbf{Counterstroke} (impose \emph{Disarray [2]}); \textbf{Screen \& Extract}.
\end{followerbox}

\begin{followerbox}{Shield-Sister Cohort}
Cap: 3 \quad Scale: I–II \\
Playbook: Combat Ally \\
Capabilities: Formation fighting, shield wall \\
Edges: Silent signals \quad Liabilities: Slow pivot \\
Strings: Clan bonds \quad Costs: Tithes; weapon upkeep \\
Tracks: Loyalty 4/6, Morale 5/6, Condition 0/3 \\
Tags: Shield Wall, Clan-Bonded, Formation Fighters
\paragraph{Orders}
\textbf{Hold the Door}; \textbf{Shield Ram}.
\end{followerbox}

\subsection{Small Strike Forces}

\begin{followerbox}{Scout-Veterans}
Cap: 3 \quad Scale: I–II \\
Playbook: Specialist \\
Capabilities: Recon, ambush, rapid deploy \\
Edges: Terrain mastery \quad Liabilities: Independent streak \\
Strings: Scout nets \quad Costs: Danger pay; kit \\
Tracks: Loyalty 3/6, Morale 4/6, Condition 0/3 \\
Tags: Scout Eyes, Rapid Deployment, Terrain Masters
\paragraph{Orders}
\textbf{Eyes Forward} (make \emph{Flank [2]}); \textbf{Night Harass} (tick \emph{Skirmish Spiral}, stay unseen \emph{or} Cohesion –1).
\end{followerbox}

\begin{followerbox}{Sapper Crew}
Cap: 3 \quad Scale: I–II \\
Playbook: Specialist \\
Capabilities: Siege, demo, field engineering \\
Edges: Structural sense \quad Liabilities: Accident risk \\
Strings: Engineer tools \quad Costs: Hazard pay; materials \\
Tracks: Loyalty 4/6, Morale 3/6, Condition 1/3 \\
Tags: Hazardous, Siege Experts, Mobile Assets
\paragraph{Orders}
\textbf{Breach} (advance \emph{Breach [4]}); \textbf{Field Fix} (restore ally Position).
\end{followerbox}

\subsection{Condotta Divisions}

\begin{followerbox}{Banner Sergeants}
Cap: 3 \quad Scale: II–III \\
Playbook: Combat Ally \\
Capabilities: Unit command, rally \\
Edges: Tactical precision \quad Liabilities: Banner pride \\
Strings: Banner rights \quad Costs: Command shares \\
Tracks: Loyalty 4/6, Morale 5/6, Condition 0/3 \\
Tags: Unit Command, Tactical Experts, Banner-Bound
\paragraph{Orders}
\textbf{Dress the Line} (stabilize \& Position +1); \textbf{Push of Pike} (advance, \emph{Rout [1]} on strong).
\end{followerbox}

\begin{followerbox}{Condotta Engineers}
Cap: 2 \quad Scale: II \\
Playbook: Specialist \\
Capabilities: Logistics, repair \\
Edges: Resource efficiency \quad Liabilities: Bureaucratic friction \\
Strings: Supply stamps \quad Costs: Admin overhead; audits \\
Tracks: Loyalty 3/6, Morale 4/6, Condition 0/3 \\
Tags: Logistics Masters, Supply Chain, Administrative
\paragraph{Orders}
\textbf{Open the Road} (reduce \emph{Supply}); \textbf{Patch \& Prep} (clear Condition or create \emph{Reserve [1]}).
\end{followerbox}

\subsection{Condotta Companies}

\begin{followerbox}{Mercenary Company (Small) — \emph{The Iron Lanterns}}
Cap: 3–4 \quad Scale: III \quad SB Resist: \(\square\) \\
Playbook: Combat Ally \\
Capabilities: Siege ops, heavy weapons \\
Edges: Bombard train \quad Liabilities: Heavy upkeep \\
Strings: Siege contracts \quad Costs: Veteran pay; trains \\
Cohorts: 20 vets, 10 support, 5 engineers \\
Tracks: Cohesion 6/8, Exposure 2/6, Resources 4/6 \\
Bandwidth: 3 orders/leg \quad Tags: Siege Specialists, Heavy Weapons
\paragraph{Orders}
\textbf{Lines \& Batteries} (Breach/Starve); \textbf{Seal the Gate} (convert \emph{Control} to tolls or safe lanes).
\end{followerbox}

\begin{followerbox}{Condotta Company (Large) — \emph{Thornwrights}}
Cap: 3 \quad Scale: IV \\
Playbook: Combat Ally \\
Capabilities: Road cutting, sapper ops, mobility \\
Edges: Bridge kits \quad Liabilities: Supply-intensive \\
Strings: Road contracts \quad Costs: Materials; fuel/forage \\
Cohorts: 40 combat, 20 engineers, 15 support \\
Tracks: Cohesion 7/8, Exposure 3/6, Resources 5/6 \\
Bandwidth: 4 orders/leg \quad Tags: Mobile Assets, Engineering Corps
\paragraph{Orders}
\textbf{March Lanes} (create \emph{Safe Lane}); \textbf{Cut \& Run} (deny road; enemy \emph{Supply –1} on strong).
\end{followerbox}

\section{War Table: The Black Banners Wargame Extension}

\subsection{Scope \& Assumptions}
This extension escalates Black Banners from company skirmishes to banner-on-banner engagements, using the core \emph{Miniatures and Tactical Layer} (ZOC, facing/flanking, special actions) and \emph{Mass Combat} scaffolding. The Keeper still adjudicates Position/Effect and DV, but units act as Followers at formation scale with clear Orders, Cohesion, and Supply interplay. % Aligns with §§8.13, 8.15. \textit{(Ref: core Combat/Conflict chapters.)}

\subsection{Table Scale \& Bases}
\begin{itemize}[leftmargin=*]
  \item \textbf{Grid}: Hex or square, as per core; declare at setup.
  \item \textbf{Base Sizes}: 
    \textit{Skirmish} (1–6 models) = Small; 
    \textit{Troop/Line} (8–16) = Medium; 
    \textit{Detachment} (20–40) = Large; 
    \textit{Wing/Battalion} (40+) = Huge.
  \item \textbf{Facing/Flank}: Use core facing; Flank = +1 die, Rear = +1 die and +1 Effect.
  \item \textbf{ZOC}: Entering enemy ZOC ends movement; leaving requires Disengage (DV 4–6) or a Boon. % Mirrors core.
\end{itemize}

\subsection{Unit Profile (Cap \& Scale)}
Each unit uses a compact profile:
\[
\textbf{Name} \;|\; \text{Type} \;|\; \text{Cap (1--5)} \;|\; \text{Scale} \;|\; \text{Speed} \;|\; \text{Tags (2)} \;|\; 
\text{Tracks: Loyalty, Morale, Condition} \;|\; \text{Costs}
\]
\paragraph{Cap (1--5)} A qualitative “tier” of training/equipment. \emph{Cap usually rises through XP milestones, training projects, and kit upgrades delivered by Quartermasters between legs.} (See \S\ref{s:cap-advancement}.)
\paragraph{Scale} Abstract size band (Small/Med/Large/Huge). Use for reach, frontage, and the \emph{Scale Edge} rule below.

\subsection{Scale Edge (quick effect dial)}
When Scale differs, apply \emph{one} of the following edge nudges (attacker chooses):
\begin{itemize}[leftmargin=*]
  \item \textbf{+1 Position} \emph{or} \textbf{+1 Effect} if attacker’s Scale $>$ defender’s.
  \item \textbf{+1 DV} against attacks if defender’s Scale $>$ attacker’s.
\end{itemize}
\textit{Design note: This keeps Scale impactful without new math bloat; it layers on top of core DV/Position.}

\subsection{Turn Structure (War Table)}
Rounds comprise the following phases:
\begin{enumerate}[leftmargin=*,label=\Roman*.]
  \item \textbf{Command} (hidden): Captains assign \emph{Orders} to each unit; reveal simultaneously.
  \item \textbf{Maneuver}: Move units (Speed, ZOC applies). Facing is chosen at end of move.
  \item \textbf{Actions}: Attack, Rally, Assist, Special, or Resolve Order effects.
  \item \textbf{End Step}: Check \emph{Cohesion}, \emph{Supply}, and event clocks; apply attrition.
\end{enumerate}

\begin{center}\small
\begin{tabular}{lll}
\toprule
\textbf{Your Unit} & \textbf{Opposition} & \textbf{Edge (crib)} \\
\midrule
Cap 0 (Levy) & Drill Troops (Cap 1) & You: Position –1 (unless bolstered) \\
Cap 1 (Trained) & Militia/Levy (Cap 0–1) & +1 Position when fresh \\
Cap 2 (Veteran) & Bandits/Levy (Cap 1–2) & +1 die on first decisive action \\
Cap 3 (Elite) & Regulars (Cap 2–3) & +1 Effect on charge/volley \\
Cap 4 (Prime) & Guard/Chosen (Cap 3–4) & Ignore 1 SB war-complication/round \\
Cap 5 (Legend) & Champions (Cap 4–5) & Once/round bump Position up one step \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Movement, Speed, Terrain, Scale.}
\textbf{Speed} is your Movement Points (MP) budget each round; you spend MP \emph{per hex} moved in the Maneuver phase.
\begin{itemize}[leftmargin=*]
  \item \textbf{Base Costs (per hex):} Open/Field = 1 MP; Road = 1 MP (ignore first Difficult modifier); Difficult (forest, rubble, marsh) = 2 MP; Very Difficult (bog, scree, shallow water) = 3 MP; Impassable = 0 MP (no entry).
  \item \textbf{Elevation:} Moving \emph{up} a level costs +1 MP; moving \emph{down} is free but may require a test if the hex is Hazardous.
  \item \textbf{Scale Modifiers (per hex):}
    \begin{itemize}
      \item \emph{Large/Huge} in \textit{Constricted or Difficult} terrain (woods, streets, rubble): +1 MP \emph{per hex}.
      \item \emph{Small} skirmish units: ignore the first +1 MP from \textit{Difficult} once per round (they flow).
      \item \emph{Road Columns:} Any Scale on a Road ignores Scale penalties while \emph{in column}; entering/leaving column costs 1 MP.
    \end{itemize}
\end{itemize}
\textit{Example:} A Large unit moving through \textbf{Difficult Forest} pays \textbf{3 MP per hex} (2 MP Difficult + 1 MP Scale). If also going \emph{up} elevation, that hex costs \textbf{4 MP}. These costs are \emph{per hex}, not a one-time surcharge.

\paragraph{Action \& Reaction Economy.}
Each unit follows a simple economy each round:
\begin{enumerate}[leftmargin=*]
  \item \textbf{Order (1)} declared in Command phase (e.g., \textit{Advance \& Hold}, \textit{Flank \& Strike}, \textit{Volley/Overwatch}, \textit{Rally \& Reform}, \textit{Sapper Work}, \textit{Logistics Action}).
  \item \textbf{Movement (Maneuver)} spends \emph{Speed} (MP) according to terrain/Scale.
  \item \textbf{Primary Action (1)} in the Actions phase, as specified by the Order (attack, rally, breach, etc.).
  \item \textbf{Reaction (up to 1/round, optional)} may trigger outside your turn when a listed trigger occurs.
\end{enumerate}
\textbf{Reactions are in addition to your primary action} but limited to \emph{one per round per unit}. Mark a small $\diamond$ beside the unit when its reaction is spent; clear all reaction markers in the End Step.

\paragraph{Common Reaction Triggers (choose one to resolve).}
\begin{itemize}[leftmargin=*]
  \item \textbf{Free Strike:} An enemy leaves your ZOC without Disengaging (DV 4–6).
  \item \textbf{Brace:} An enemy \emph{enters} your front arc from \(\ge\) 2 hexes away; gain Position +1 versus that charge/attack.
  \item \textbf{Overwatch Fire:} If your unit set \textit{Volley/Overwatch} this round, make one prepared attack when the first enemy crosses your lane.
  \item \textbf{Counterpush:} If an enemy resolves a \textit{Shock Push} into your hex, test to Shove 1 hex back on a hit (opposed).
\end{itemize}
\textit{Notes:} (1) \textit{Skirmish Screen} grants a \emph{free Disengage} during your move; it is \emph{not} a reaction. (2) Orders that set a posture (e.g., \textit{Volley/Overwatch}) \emph{consume your action}, and the triggered shot uses your one reaction for the round. (3) Some tags (e.g., \textit{Line-Drilled}) may grant a \emph{bonus reaction} 1/round; if so, it will be stated on the unit.

\paragraph{Cheatline:}
\[
\textit{Order (1)} \;\Rightarrow\; \textit{Move (MP)} \;\Rightarrow\; \textit{Primary Action (1)} \;\;\;\; \big|\;\; \textit{Reaction: max 1/round when triggered}
\]

\subsection{Orders Economy (choose one per unit/round)}
\begin{description}[leftmargin=*,style=nextline]
  \item[Advance \& Hold] Move up to Speed, then Guard. Good for line stabilization.
  \item[Flank \& Strike] Move; if you end in a flank/rear arc, gain +1 die on one attack.
  \item[Skirmish Screen] Enter/exit ZOC freely once (auto-Disengage), but attacks are Limited.
  \item[Shock Push] Attempt a Tackle/Shove (DV 4–6); on success, Push 1–2 hexes and degrade foe Position.
  \item[Volley/Overwatch] Gain Aim benefits (+1 die or +1 Effect), threaten a lane as a Reaction.
  \item[Rally \& Reform] Test \emph{Morale}; on success, clear 1 \emph{Morale} and 1 \emph{Condition}.
  \item[Sapper Work] Place/clear obstacles, breach, or entrench; set a Terrain Tag for your hex/ring.
  \item[Logistics Action] Reduce Company \emph{Supply} by 1 to remove 1 \emph{Condition} from this unit.
\end{description}

\subsection{Cap vs. Typical Foes (Keeper crib)}
\begin{center}\small
\begin{tabular}{lll}
\toprule
\textbf{Your Unit} & \textbf{Opposition} & \textbf{Edge} \\
\midrule
Cap 1 (Trained) & Militia (Cap 0--1) & +1 Position when fresh \\
Cap 2 (Veteran) & Bandits/Levy (Cap 1--2) & +1 die on first decisive action \\
Cap 3 (Elite) & Regulars (Cap 2--3) & +1 Effect on charge/volley \\
Cap 4 (Prime) & Guard/Chosen (Cap 3--4) & Resist 1 SB war-complication/round \\
Cap 5 (Legend) & Champions (Cap 4--5) & Once/round, bump Position up one step \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Followers \& Companies at War}
\paragraph{Tracks that Talk.} Individual \emph{Loyalty/Morale/Condition} bleed upward: at the end of any round where $\ge 2$ units are at \emph{Morale} $\le 2$ or \emph{Condition} $\ge 2$, tick Company \emph{Cohesion} +1. When Company \emph{Cohesion} drops a step (e.g., 4/6 $\rightarrow$ 5/6), all units take \emph{Position} –1 until a Rally succeeds. % Ties Followers to Company state.
\paragraph{Costs with Teeth.} If a unit’s \emph{Costs} go unpaid this leg, it immediately suffers \emph{Loyalty} –1 \emph{and} starts a \emph{Desertion [3]} clock; on fill, remove the unit or convert it to \emph{Bannerless} (hostile neutral).
\paragraph{Veteran Resilience.} Veteran/Elite units may \emph{ignore one SB war-complication per battle} (once)—or convert it into \emph{Condition +1} instead.

\subsection{Cohesion, Payday, Supply (Company Sync)}
\begin{description}[leftmargin=*,style=nextline]
  \item[Payday (4)] Missed payments or broken writs tick this. On fill: \emph{Loyalty} –1 to all units; begin \emph{Pay Raid [4]} as soldiers self-requisition.
  \item[Supply (4)] Weather, hard marches, or requisition burn. On fill: all \emph{Logistics Actions} fail this round; ranged units lose \emph{Volley}.
  \item[Cohesion (6)] Losses, betrayals, poor command. Each step past 4/6 imposes army-wide Position –1 until \emph{Rally \& Reform} succeeds twice.
\end{description}

\subsection{Terrain \& Weather (quick plugs)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Difficult}: 2 MP/hex; \textbf{Elevation}: +1 DV from below; \textbf{Cover}: Position up a step.
  \item \textbf{Storm}: DV –1 for movement tests; \textbf{Snow}: Position –1 unless \textit{Cold-Drilled}.
\end{itemize}
% Mirrors core terrain and weather dials.

\subsection{Orders—“When to Use” Quick Guide}
\begin{itemize}[leftmargin=*]
  \item \textbf{I need to pin their center} $\rightarrow$ Advance \& Hold (then Guard).
  \item \textbf{I need a decisive break} $\rightarrow$ Shock Push into Rear/Flank.
  \item \textbf{Their light troops harass us} $\rightarrow$ Skirmish Screen.
  \item \textbf{We’re wavering} $\rightarrow$ Rally \& Reform (twice if Cohesion is slipping).
  \item \textbf{We can grind them} $\rightarrow$ Volley/Overwatch to tax approaches.
  \item \textbf{Fort ahead} $\rightarrow$ Sapper Work to set breach/climb tags.
\end{itemize}

\subsection{Active Orders Box (per unit)}
\begin{center}\small
\begin{tabular}{p{0.92\linewidth}}
\toprule
\textbf{Current Deployment}: \_\_\_\_\_ \quad \textbf{Active Order}: \_\_\_\_\_ \quad \textbf{Objective}: \_\_\_\_\_ \\
\textbf{Status}: Loyalty \_\_/6 \; Morale \_\_/6 \; Condition \_\_/3 \;\;|\;\; \textbf{Costs Due}: \_\_\_ \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Cap Advancement \& Projects}\label{s:cap-advancement}
\begin{itemize}[leftmargin=*]
  \item \textbf{Training Project} (\emph{Company Phase}): Spend coin/scrip; complete \emph{Drill [6]}; raise one unit’s \emph{Cap} by +1 (max 5). On completion, mark \emph{Supply} +1 (ammo/equipment burn).
  \item \textbf{Kit Upgrade}: Attach a Diamond-tier asset (engine, plate, chassis); gain a situational Tag for the campaign (\emph{Siege-Drilled}, \emph{Cold-Drilled}, \emph{River-Rated}).
  \item \textbf{Veteran Cross-posting}: Move a veteran cadre into a green unit; both roll \emph{Morale}. On success, mentees gain temporary +1 die on their next Rally or Advance.
\end{itemize}

\subsection{SB Menu for War (Keeper)}
\begin{itemize}[leftmargin=*]
  \item \textbf{False Orders} (colors flipped at a ford)
  \item \textbf{Ally Turns Sides} (Favor Debt called)
  \item \textbf{Cut Lines} (Supply +1, Cohesion +1)
  \item \textbf{Camp-Fever} (disease outbreak; Condition +1 unless Chaplains/Leeches intervene)
  \item \textbf{Gear Shear} (siege tool failure; lose Sapper Work this round)
  \item \textbf{Authority Edict} (political interference; Position –1 in parley or movement)
\end{itemize}

\subsection{Setups \& Diamonds}
\begin{itemize}[leftmargin=*]
  \item \textbf{River Prefect Stamps}: Crossing DV –1; one forced march ignores Difficult once.
  \item \textbf{Broker’s Favor}: Parley DV –1 on contract scenes; on miss, gain a Black Mark.
  \item \textbf{Bell-Tokens}: Safer prisoner handling; cancel one \emph{Ally Turns Sides}.
\end{itemize}

\subsection{Worked Micro-Scenario}
\paragraph{“Banner Swap in the Wind.”} A rival plants \emph{False Orders}. Roll \emph{Allegiances in Flux} at start. Your Vanguard (Cap 3, Medium, Tags: \emph{Line-Drilled}, \emph{Cold-Drilled}) declares \emph{Advance \& Hold}; your Lights (Cap 2, Small, \emph{Skirmish}) declare \emph{Skirmish Screen}. Enemy hits your flank; you answer with \emph{Shock Push} from Reserve (Cap 3, Large \emph{Cavalry}). If two of your units end the round with \emph{Morale} $\le 2$, tick Company \emph{Cohesion} +1 and apply army-wide Position –1 until double Rally succeeds.

\subsection{Reference Tabs}
\begin{itemize}[leftmargin=*]
  \item \textbf{Facing/Flank}: +1 die (Flank), +1 die and +1 Effect (Rear).
  \item \textbf{ZOC}: Enter stops, exit needs Disengage (DV 4–6) or Boon.
  \item \textbf{Terrain}: Difficult = 2 MP/hex; Elevation = +1 DV from below; Cover raises Position.
  \item \textbf{Reactions}: Each unit 1/round (Free Strike or Shove 1 on ZOC leave).
\end{itemize}

\end{document}
