% way_of_silk.tex
\documentclass[11pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{a5paper, margin=1in}
\usepackage{setspace}
\onehalfspacing
\usepackage{microtype}
\usepackage{ebgaramond}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{array}
\usepackage{tabu}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=black, citecolor=black, urlcolor=gray}

\titleformat{\chapter}[hang] 
{\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{0pt}{-20pt}{20pt}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[RE]{\textit{\leftmark}}
\fancyhead[LO]{\textit{\rightmark}}
\renewcommand{\headrulewidth}{0.4pt}

\newenvironment{lorebox}{\begin{quote}\small\itshape}{\end{quote}}

\begin{document}

\frontmatter
\begin{titlepage}
\centering
\vspace*{2cm}
{\huge\bfseries The Way of Silk\par}
\vspace{0.5cm}
{\Large A Fate's Edge Expansion\par}
\vspace{2cm}
{\large Caravans, Cultures, and the Road That Binds the East\par}
\vspace{1.5cm}
\vfill
\end{titlepage}

\tableofcontents

\mainmatter

\chapter*{Preface: The Road Is a River}
\addcontentsline{toc}{chapter}{Preface: The Road Is a River}

The maps of the empire show the Way of Silk as a single line -- a thread from the Amaranthine to the Eastern Ocean. Those who walk it know the truth: the Way is a river. It has shallows and rapids, tributaries and deltas. Its current carries not only silk and spice but also gods, spies, and the debts of a dozen kingdoms.

This expansion is not a traveller's gazetteer of “exotic” lands. It is a ledger of the road itself. The peoples who live along the Way -- the Fhara, the Sidhi, the Pereshi, the Kuvani, the Tulkani, and the divided Dhaharan -- are not backdrops. They are the toll‑keepers, the ferrymen, and the witnesses to every broken promise. Their etiquette is not decoration; it is a negotiation. Their patrons are not distant gods but forces that manifest in the creak of the cart and the smell of rain on dry dust.

We have drawn from the stories of the great overland routes, the caravanserais of the western deserts, the mountain passes of the eastern steppes, and the oasis courts of the southern coasts. But we have not assembled a museum. The Dhaharan here is split between two administrations -- the mountain viceroys and the coastal admiralty -- because trade routes create contradictions, not harmonies. The Tulkani spread the cult of Ikasha not through temples but through knots tied in wagon cords. And the amulet Kalrantha, a shard of glass and brass, is a world‑shaking artifact only in the fear it inspires; its true power was sealed from the moment it was stolen.

When you play along the Way of Silk, you will haggle over water‑rights, not holy relics. You will arbitrate disputes between a Fhara water‑clerk and a Kuvani remount master. You will sleep in caravanserais where the walls remember the last three murders. And if you are wise, you will listen when a Tulkani storyteller whispers: \textit{“The road does not forgive a debt. It only reschedules it.”}

The ninth measure -- the silence, the empty space, the unsaid name -- still applies.

\clearpage
\chapter{The Road Itself}

\section{Geography of the Way}

The Way of Silk is not a single route but a braid of paths that shift with season, war, and the price of fodder.

\begin{itemize}
    \item \textbf{The Northern Way:} The western steppe (home to Vili and Ostrikari humans, and Ykrul clans) \(\to\) Aeler passes \(\to\) Sihai frontier. Cold, hard, and fast. Ykrul remount posts are spaced at a day’s ride; the Aeler keep the passes open with breath‑books and toll‑ledgers.
    \item \textbf{The Central Oasis Route:} Fhara wells (on the great southern peninsula) \(\to\) Pereshi highlands \(\to\) Kuvani summer pastures \(\to\) Dhaharan uplands. Slow and fraught with water politics. Each well has a keeper, each keeper a price.
    \item \textbf{The Monsoon Coast:} Kahfagian ports \(\to\) Ashaani river plains \(\to\) Ayokha delta. Wet and dangerous. The Ashaani press‑gangs and the Kahfagian privateers share a love of false flags.
    \item \textbf{The Himadri Spur:} Aeler mountain passes connecting the Dhaharan plateau to the northern plains. A single pass is a kingdom; to control it is to tax half the silk trade.
\end{itemize}

\section{Caravan Procedures}

Use the travel clocks and leg structure from \textit{Caravans: Way of Silk}. Add the following intent dial for the Way:

\begin{itemize}
    \item \textbf{Speed:} press camels, risk lameness.
    \item \textbf{Stealth:} avoid bandits and inspectors.
    \item \textbf{Show:} fly banners, gain respect, attract spies.
    \item \textbf{Tribute:} travel as a bonded caravan of a patron; gain protection but pay a portion of all profits.
\end{itemize}

\subsection*{New Hazards of the Way}
\begin{itemize}
    \item \textbf{Dry well:} The oasis keeper is dead. Lose 1 Supply and roll on the Faction table.
    \item \textbf{Forged seal:} Your papers are false; the local governor demands a bribe or a favour.
    \item \textbf{Ghost caravan:} A train of skeletal camels passes at midnight. The next day, one of your beasts becomes a spirit for an hour.
    \item \textbf{Debt collector:} Someone recognises a Tulkani vow‑cord you carry. They demand payment in story or a favour to be named later.
\end{itemize}

\clearpage
\chapter{Peoples of the Road}

\section{Fhara -- Oasis Caravaneers}
\subsection*{Homeland}
The Fhara dwell on a great, sun‑baked peninsula in the south‑west, where scattered oases and dry riverbeds connect the coast to the interior. Their cities are built around wells and date groves, and their caravans are the arteries of the spice trade.

\subsection*{Pillars}
Water‑law, tent courts, coffee ritual, lamp‑lit contracts.

\subsection*{Voice}
Fhara speak with a dry precision, like a coin chipped to test its purity. They negotiate in couplets -- a verse of praise, a verse of price.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Offer coffee first} -- Position +1 for trade talks.
    \item \textbf{Speak in three cups} -- first cup introduces, second cup bargains, third cup seals. Rushing the second is an insult (GM gains 1 SB).
    \item \textbf{Never refuse a date} -- the fruit is sacred to their well‑spirits. (Declining ends the conversation.)
\end{itemize}

\subsection*{Strings}
Water‑share tablet, caravan court writ, escort oath, lamp‑sealed contract.

\section{Sidhi -- Coastal Cousins of Ashaan}
\subsection*{Homeland}
The Sidhi are a minority people whose ancestral homeland is Galanina, a rugged coast of limestone hills and ancient ports far to the west. Many now live as traders and sailors in Oshiiran ports, while others were taken as slaves to Ashaan and still others have found refuge in the Brass Coast of Dhahara. They are fierce, proud, and never forget a debt.

\subsection*{Pillars}
Sea trade, lighthouse courts, mixed ancestry (Vilikari and Ashaani), fierce independence.

\subsection*{Voice}
Sidhi speak with a musical lilt, their grammar a creole of Ashaani and Vilikari. They are direct, suspicious of strangers, and quick to laugh.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Light a lamp before docking} -- Position +1 in Sidhi ports.
    \item \textbf{Never ask about ancestry} -- the Ashaani slavery raids are a fresh wound. (Breaking this flips Audience from Warm to Cold.)
\end{itemize}

\subsection*{Strings}
Lighthouse charter (right to harbour), port‑weight (trade advantage), freedman’s seal (proof of emancipation from Ashaan).

\section{Pereshi -- Inland Highlanders}
\subsection*{Homeland}
The Pereshi inhabit the high plateaus and mountain valleys that lie between the Fhara peninsula and the eastern steppes. Their ancient empire fell long ago, leaving a patchwork of petty kingdoms, fire temples, and caravanserai lords. They are the scholars, magi, and bureaucrats of the Silk Way.

\subsection*{Pillars}
Ancient libraries, garden shrines, caravan tariffs, messenger roads.

\subsection*{Voice}
Pereshi speech is formal and layered, filled with honourifics and allusions to forgotten poets. A question may be disguised as a proverb; an insult as a compliment.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Present a red rose before a petition} -- Position +1 for official audiences.
    \item \textbf{Never step on a fallen book} -- written words carry the weight of law. (Violation starts a 2‑segment “Disrespect” clock.)
    \item \textbf{Recite a verse when crossing a bridge} -- the spirits of the water expect it. (Failure to do so imposes Desperate Position on the next river hazard.)
\end{itemize}

\subsection*{Strings}
Fire‑temple token, caravan tariff receipt, royal seal (on loan from a minor prince), messenger tablet.

\section{Kuvani -- Steppe Hosts}
\subsection*{Homeland}
The Kuvani range across the eastern steppe, a vast grassland east of the Valewood. They are a relatively new people in the region, their clan confederations only recently rising to prominence. They have not yet become the ruling class, but their horsemanship and remount stations make them indispensable to any caravan that crosses the open plain.

\subsection*{Pillars}
Remount lines, wind‑bands, felt tents, kurgan oaths.

\subsection*{Voice}
Kuvani speak as if the wind were an impatient audience -- short, quick, and punctuated by gestures toward the horizon.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Praise the horse before the rider} -- DV --1 for all negotiations.
    \item \textbf{Ride at the host’s pace} -- never pass the guide. (Breaking this starts a 4‑clock “Insult to the Herd”.)
\end{itemize}

\subsection*{Strings}
Remount chain mark, grazing compact, sky‑road song, kurgan oath (sealed with horsehair).

\section{Tulkani -- Itinerant Courts}
\subsection*{Homeland}
The Tulkani have no fixed homeland; they are wanderers who follow the seasons and the trade routes. Their wagons are their hearths, and their circles are their courts. They are welcome everywhere and trusted nowhere -- a position they wear with wry humour.

\subsection*{Pillars}
Caravan courts, tinkers, musicians, neutrality customs. Agents of Ikasha often walk among them.

\subsection*{Voice}
Tulkani speak in riddles and stories. A bargain is a parable; a threat is a lullaby. They never say “no” -- only “that road is closed for repairs”.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Tie a knot for each truth} -- you may cancel one SB by adding a truth (a secret, a confession) to your vow‑cord.
    \item \textbf{Never sleep outside the circle} -- the wagon ring is sanctuary. (Leaving it at night gives the GM 1 SB to use for “something follows”.)
\end{itemize}

\subsection*{Strings}
Vow‑bead, safe‑haven oath (neutral camp), caravan judge’s writ, dream‑cord (gift from Ikasha).

\section{Dhahara -- The Split Land (Preview)}

\subsection*{The Northern Mountain Satrapies}
Aeler viceroys rule the high valleys, collecting taxes in silver and tea. Their courts blend mountain custom with lowland ceremony; elephant stables and fire‑shrines stand side by side. The local Aeler have adopted Dhaharan dress and speak a creole of their own tongue. They are loyal to the Mountain Viceroy, who answers to the Aeler High King.

\subsection*{The Southern Brass Coast}
Administered by the Kahfagian Admiralty, the Brass Coast is a string of ports connected by rope‑ways and spice warehouses. The Admiralty treats Dhahara as a “client state” – profitable, exotic, and restive. The Sidhi are most numerous here, and they keep their own lighthouse courts alongside the Admiralty’s maritime tribunals.

\subsection*{Two Loyalties Mechanic}
Every Dhaharan character or faction has two tracks: \textbf{Loyalty to the Mountain} and \textbf{Loyalty to the Admiralty}. Adventures often force a choice between them; losing favour with one may gain favour with the other.

\subsection*{Adventure Seed: The Viceroy’s Seal}
The Mountain Viceroy’s seal has been stolen and sold to a Brass Coast merchant. Retrieve it before the Admiralty uses it to forge a treaty that would make the north a colony. The party is hired by a Sidhi dock‑master – who secretly works for the Breath‑less (Ashaani resistance) and has his own plans for the seal.

\subsection*{Full expansion planned: \textit{Dhahara: Throne of Brass}.}

\clearpage
\chapter{Patrons of the Way}

\section{Regional Faces of Core Patrons}

\subsection{The Traveler – Lord of the Forked Road (Eastern variant)}
Known as \textbf{Jie Lu}, the Road that Asks. His shrines are empty crosses at crossroads; his ceremonies require walking barefoot from one gate to another.

\textbf{Etiquette:} Before speaking of profit, name a place you have never been. (Cancels one social SB about greed.)

\textbf{Rite -- The Unlatched Gate:} Once per journey, declare a road closed; it becomes passable for your party only.

\subsection{Ikasha – She Who Sleeps in the Caravan (Tulkani variant)}
Among the Tulkani, Ikasha is the dream inside the wagon. She speaks through creaking wheels and the scent of rain on canvas. Her cult is spread through story‑ledgers -- knots tied in a cord, each knot a secret told to a stranger.

\textbf{Etiquette:} At dusk, whisper a fear into a cup of water. Pour it onto the road. (GM gains 1 SB, but you gain +1 Position on your next Stealth roll.)

\textbf{Rite -- The Wagon Circle:} When you camp in a circle, no spirit may cross the line of wheels without an invitation.

\subsection{Mykkiel – The Judge of Galanina}
In Ashaan, Mykkiel is not a distant arbiter but a direct antagonist to the Veiled Sisters. His followers are the secret courts of the Breath‑less, the judges who try the Sisters in absentia. Mykkiel demands that every oath be written and witnessed; the Sisters rule through unspoken terror. It is a cold war of ink versus silence.

\textbf{Regional Rite -- The Unsealed Testimony:} Once per session, you may compel a spirit to speak truth for one question, but the spirit’s former master learns your face. (Mark +2 Obligation)

\subsection{Aveh – The Dust‑Rider (Dhaharan variant)}
Here, Aveh is a djinn -- a spirit of the sandstorm, the liberator of slaves, the patron of those who flee without paying. She appears as a rider without a face, offering an exit from any prison or contract. Her price is always the same: you must never close a door behind you.

\textbf{Etiquette:} When you flee, leave a coin at the threshold. (The GM may convert the first SB of pursuit into a “missed footing” complication instead of harm.)

\textbf{Rite -- The Unlatched Chain:} Break one physical bond (shackle, rope, locked door) as a free action. (Mark +1 Obligation; the broken object may later reform.)

\section{New Patrons of the Eastern Road}

\subsection{Nasima – The Oasis That Moves}
\paragraph{Domain:} Springs, mirages, mercy of water.
\paragraph{Favoured Deeds:} Sharing water, digging a well, guiding the lost without payment.
\paragraph{Taboos:} Polluting a spring, selling water to the dying, lying about a well.
\paragraph{Low Rite -- The Mirage Cup (4 XP):} Pour water onto sand; know the nearest clean water for a scene. (Mark +1 Obligation)
\paragraph{Standard Rite -- Nasima’s Blessing (8 XP):} A waterskin becomes bottomless for a day. (Mark +2 Obligation)
\paragraph{High Rite -- The Oasis Pact (14 XP):} Create a temporary well that lasts a season. (Mark +3 Obligation; may attract unwanted spirits)

\subsection{Div – The Whisper in the Ruin}
\paragraph{Domain:} Ruins, forgotten knowledge, echoes of empire.
\paragraph{Favoured Deeds:} Uncovering forgotten truth, destroying false chronicles, letting ruins rest.
\paragraph{Taboos:} Using a relic for personal gain, selling a historical secret to a tyrant.
\paragraph{Low Rite -- Echo of the Lost (4 XP):} Hear last words spoken in a ruin. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Div’s Accounting (8 XP):} See invisible traces of past violence. (Mark +2 Fatigue)
\paragraph{High Rite -- The Unsealed Archive (14 XP):} Open a sealed door or memory; GM reveals one hidden truth. (Mark +3 Corruption)

\subsection{Zun – The Shepherd of Scales}
\paragraph{Domain:} Price, balance, fair exchange.
\paragraph{Favoured Deeds:} Correcting false weights, mediating trade disputes.
\paragraph{Taboos:} Using counterfeit coin, undervaluing labour.
\paragraph{Low Rite -- Weight of Truth (4 XP):} Know if an object was obtained dishonestly. (Mark +1 Obligation)
\paragraph{Standard Rite -- Zun’s Ledger (8 XP):} Guarantee 10\% profit on a trade, but on a miss the profit turns to salt. (Mark +2 Obligation)
\paragraph{High Rite -- The Scale of Judgement (14 XP):} Mirror harm between two opponents for one day. (Mark +3 Obligation)

\subsection{Al‑Qibla – The Stone That Looks West}
\paragraph{Domain:} Prayer, direction, geometry of faith.
\paragraph{Favoured Deeds:} Building a shrine, drawing accurate maps, praying in a foreign land.
\paragraph{Taboos:} Defacing a place of worship, selling misleading maps.
\paragraph{Low Rite -- The True Bearing (4 XP):} Always know direction to nearest shrine of your faith. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Stone’s Sanctuary (8 XP):} Create a circle that grants +1 die to resist fear or charm. (Mark +2 Fatigue; you may not leave the circle.)
\paragraph{High Rite -- The Unerring Caravan (14 XP):} Cannot be lost for one leg; you must walk at the front and never look back. (Mark +3 Obligation)

\subsection{Peri – The Fire That Hides}
\paragraph{Domain:} Caves, hidden springs, the spark waiting.
\paragraph{Favoured Deeds:} Creating hiding places, burying secrets, escaping capture.
\paragraph{Taboos:} Revealing a cache you were trusted with, using a hiding place to ambush a guest.
\paragraph{Low Rite -- The Hidden Door (4 XP):} One touched doorway becomes invisible to those without the password. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Peri’s Cache (8 XP):} Bury an object; it cannot be found for one month. (Mark +2 Fatigue)
\paragraph{High Rite -- The Cave of Sleep (14 XP):} Create a temporary pocket that can hold six people for one night. (Mark +3 Obligation)

\subsection{Khidr – The Green Man of the Ford}
\paragraph{Domain:} Water crossings, thresholds of the river, rescue.
\paragraph{Favoured Deeds:} Saving drowning people, maintaining fords, helping travellers cross dangerous water.
\paragraph{Taboos:} Asking payment from a corpse, letting a bridge collapse from neglect.
\paragraph{Low Rite -- The Hand of the Ford (4 XP):} Pull an ally from water as a reaction. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Khidr’s Path (8 XP):} Allies gain +1 die to any roll crossing water. (Mark +2 Fatigue)
\paragraph{High Rite -- The Second Crossing (14 XP):} Once per campaign, revive an ally who died within sight of water – but they forget the last year. (Mark +3 Corruption)

\subsection{Daeva – The Debt That Screams}
\paragraph{Domain:} Unpaid vows, broken promises, interest that accrues.
\paragraph{Favoured Deeds:} Collecting a bad‑faith debt, destroying a false contract, exposing a usurer.
\paragraph{Taboos:} Extending an unpayable loan, forgiving a debt without witness.
\paragraph{Low Rite -- The Tally of Silence (4 XP):} Know if a contract will be broken. (Mark +1 Corruption)
\paragraph{Standard Rite -- Daeva’s Mark (8 XP):} Curse an oath‑breaker with –1 die to all rolls until amends. (Mark +2 Corruption; may rebound.)
\paragraph{High Rite -- The Ledger of Wails (14 XP):} Haunt an oath‑breaker’s bloodline – they cannot sleep, fatigue track doubles. (Mark +3 Corruption; once per campaign.)

\clearpage
\chapter{The Amulet of Kalrantha}
\label{chap:kalrantha}

\subsection*{What It Is}
A small, tear‑shaped amulet of green glass, set in a brass frame that never tarnishes. Warm to the touch when someone nearby is about to make a decision. To those who can sense such things, it hums with the element of \textbf{Fate} – not the fate of prophecies or chosen ones, but the cold, indifferent gravity of consequence. Fate does not guide; it \textit{pulls}. Like a stone falling toward the ground, the amulet does not care whether the landing is soft or hard. It only knows that the fall must end.

\subsection*{Its History (As Told by the Tulkani)}
\begin{lorebox}
“Long ago, a prince of Dhahara stole this from a temple where no lamps were lit. He thought it would let him choose his own fortune. He wore it while he conquered a kingdom. The desert ate that kingdom. The prince’s name is not written anywhere, but his shadow is still seen in the Siq at dawn.”
\end{lorebox}

\subsection*{The Truth (GM Only)}
The amulet was created by the priests of that temple as a \textit{trap} – a vessel for a demon of indecision. The prince, in his arrogance, released the demon into the world, but the amulet itself was designed to contain it. The disaster that swallowed his kingdom was not a curse; it was the demon using the prince as a puppet. Now the amulet is empty – or almost empty. A fragment of the demon remains, dormant, waiting for a wielder to make a fateful choice. The amulet does not grant wishes; it \textit{offers suggestions}. The horror is that the characters will never know if the suggestion was their own or the amulet’s.

But the amulet’s power is not the demon’s alone. It is woven from the raw element of \textbf{Fate} – the same force that makes oaths bind and thresholds remember. Fate is anti‑magic: it does not create, it does not transform; it only \textit{ensures}. A spell that tries to alter destiny will sputter and die in the amulet’s presence. A lie that would change the future will ring hollow. The amulet pulls events toward their most inevitable conclusion, whether that conclusion is a triumph or a grave.

\subsection*{Game Mechanics}

The amulet is a **Tier V artifact** with the following properties:

\begin{itemize}
    \item \textbf{Fate’s Gravity:} Once per session, the wearer may reroll a single die after seeing the result. This is a mundane luck effect – the amulet’s smallest gift.
    \item \textbf{Anti‑Magic Aura (Near):} Any spell or magical effect that directly alters probability, fate, or causality (e.g., a Rite that forces a reroll, a blessing that guarantees success, or a curse that imposes automatic failure) suffers \textbf{+2 DV} within Near range of the amulet. The amulet’s wearer is immune to this penalty – they are the anchor.
    \item \textbf{Story Beats Become Gravity:} This is the amulet’s true, unsettling power. When the wearer rolls one or more \textbf{1s} (generating Story Beats for the GM), those Story Beats are not added to the GM’s pool. Instead, they are \textit{transferred} to the \textbf{next player character who acts}. That player gains those Story Beats as if they had rolled them – but the beats manifest as a pull toward fate. The GM must spend them on complications that affect that character’s action, and those complications must be \textit{inevitable} in tone (e.g., “the floor gives way,” “your sword strikes a hidden chain,” “you recognise your own face in a reflection – and it is weeping”).
    \item \textbf{Mechanical Summary:}
    \begin{itemize}
        \item Wearer rolls 1s → those SB are not given to GM.
        \item Instead, the next PC to act gains those SB as a separate pool.
        \item The GM spends those SB on complications for that PC’s action only.
        \item Any unspent SB from this transfer vanish after that action.
    \end{itemize}
    \item \textbf{Obligation and the Demon Fragment:} The amulet itself does not generate Obligation – the demon fragment does. Whenever the wearer uses the reroll ability or benefits from the anti‑magic aura in a scene, mark \textbf{+1 Obligation} to the amulet (treat as a Terrestrial Patron with Capacity = Spirit). When Obligation exceeds Capacity, the fragment stirs. The wearer has nightmares of falling. The GM gains 1 SB per excess segment. When Obligation reaches double Capacity, the fragment manifests as a voice whispering a single suggestion: “Take the left path,” “Trust the merchant,” “Draw your sword.” The wearer may ignore it – but doing so immediately clears all Obligation and the amulet becomes inert for one session.
\end{itemize}

\subsection*{Anti‑Magic and Fate – Examples}
\begin{itemize}
    \item A Runkeeper tries to invoke a Rite that would guarantee an ally’s next attack hits. Near the amulet, the DV increases by +2. The amulet’s wearer (if not the caster) is unaffected; fate favours the anchor.
    \item A Cantor sings a song to reroll a failed navigation check. Within the amulet’s aura, the song falters – the DV increase applies. The Cantor feels a strange certainty that the original roll \textit{was} the correct path, even if it led to a dead end.
    \item A Free Caster attempts a spell to turn an enemy’s blade aside. The amulet does not block the spell – it only resists changes to probability. The blade is still turned, but a nearby lantern falls, or a loose stone shifts underfoot. Fate balances the scales.
\end{itemize}

\subsection*{Story Beat Transfer in Play – Example}
\begin{enumerate}
    \item Kael (wearer of the amulet) rolls to pick a lock. He rolls two 1s. Normally the GM would gain 2 SB. Instead, the SB are not added to the GM’s pool. They are held.
    \item Next, Lyra attempts to sneak past a guard. The GM announces: “You have 2 SB from the amulet to spend on Lyra’s action.”
    \item Lyra rolls a Partial success. The GM spends 1 SB: “As you move, a loose board creaks – not loud enough to alert the guard, but enough that your cloak snags on a nail. Mark 1 Fatigue.” The second SB is banked for later in the same action, but not used.
    \item The GM cannot spend those SB on anyone else. After Lyra’s action, any unspent SB vanish.
\end{enumerate}

\subsection*{Adventure Hook: The Empty Vessel}
A Breath‑less agent in Ashaan claims the amulet can bind the Veiled Sisters’ greatest spirit. She is wrong – it can only hold a fragment – but the Sisters believe her. The party must reach the amulet before the Sisters do, then decide whether to use it, destroy it, or hide it again.

\subsection*{GM Guidance}
The amulet of Kalrantha is designed to create a sense of creeping inevitability. The players should feel that wearing it is a burden, not a gift. The anti‑magic aura will frustrate some spellcasters, but the story beat transfer creates a unique tactical puzzle: the wearer’s failures become another player’s complications. This encourages the party to plan around who wears the amulet and when. A wise party might give it to the character who rolls the most dice (and thus the most 1s) – but that character will also suffer the anti‑magic penalties when near their own allies.

The demon fragment should remain a mystery for most of the campaign. Its voice is a whisper, never a command. Its suggestions are always plausible, never obviously harmful. The horror is not that the amulet controls you – it is that you will never know if the choice was yours.

\begin{lorebox}
“Fate is not a path. It is the weight of the ground when you stop falling.”
--- \textit{Dusana of the Raven Road, on the amulet Kalrantha}
\end{lorebox}

\clearpage
\chapter{Adventure Framework}

\section{Adventure Seeds (Ten Drop‑Ins)}
\begin{enumerate}
    \item \textbf{The Vanishing Oasis:} A well the party used last year is gone. The Fhara water‑clerk is dead. Find what erased it – and who is poisoning the remaining wells.
    \item \textbf{The Tulkani Arbitration:} Two caravan masters invoke the Tulkani oath‑ring to settle a dispute. The party is the witness. The truth is that both are corrupt; a third party planted the evidence.
    \item \textbf{The Forged Seal of Sihai:} A merchant’s Sidhi seal is a forgery. The party must find the forger – who is a disgraced official hiding in the Brass Coast.
    \item \textbf{The Ghost of the Great Caravan:} Once a year, a spectral caravan crosses the desert. The party follows it to a buried ruin where an old debt must be paid in memory.
    \item \textbf{The Coffee Challenge:} A Taharkan prince offers a fortune to race his horse against an Oshiiran steed. The loser’s kingdom forfeits a pass.
    \item \textbf{The Kiln That Dreams:} A Kuvani remount station has a pottery kiln that produces perfect horse‑charm jars – but the spirits inside are children. The maker is a desperate potter who lost a child to Ashaani slavers.
    \item \textbf{The Lighthouse Court:} A Sidhi lighthouse captain is accused of smuggling. The party must defend her in a court of port‑masters – or help her escape if she really is guilty.
    \item \textbf{The Brass Gate Bribe:} A high official of the Brass Gate is taking bribes from Ashaan. The party is hired to expose him – but the evidence is in a locked archive that Mykkiel judges as “inadmissible” without a witness.
    \item \textbf{The Road That Sleeps:} A stretch of the Way has been cursed. Every traveller who sleeps there dreams of drowning. The cause is an old contract with Aveh that was never closed.
    \item \textbf{The Ninth Cord:} A Tulkani storyteller offers to tell the party a secret for free – a ninth story that never ends. If they listen, they owe her a favour to be named later.
\end{enumerate}

\section{Showpiece: The Ghost of the Great Caravan}
A multi‑session mystery across three legs. The party is hired to escort a Tulkani wise‑woman to a ruin where, according to legend, the first silk caravan was destroyed by a sandstorm. The spirits of that caravan re‑enact their death every year. The wise‑woman wants to steal the amulet Kalrantha from the ruins; she believes it can speak to the dead. In truth, the amulet is a decoy; the real treasure is a map that reveals a forgotten pass through the mountains. The map is written in the language of Mykkiel’s old court – and the Sisters will kill to keep it hidden.

\clearpage
\chapter{GM Toolkit}

\section{Travel Hazards Table (d12)}
\begin{enumerate}
    \item Dry well – lose 1 Supply.
    \item Bandit embassy – they offer safe passage for a toll. Refuse and start a Pursuit clock.
    \item Ghost caravan – pass peacefully, but one beast becomes a spirit for an hour.
    \item Sandstorm – Navigator test DV 4 or lose 2 segments of progress.
    \item Customs inspection – roll Papers or start a 4‑clock Audit.
    \item Oasis court – a dispute over water rights; mediate or be delayed.
    \item Elephant stampede – test Stealth or take Harm 1.
    \item Mirage – a false well; lose 1 Supply and mark Fatigue.
    \item Debt collector – someone recognises your Tulkani vow‑cord; pay in story or favour.
    \item Ashfall – from distant volcano; all actions Desperate due to low visibility.
    \item Singing dunes – the sand hums. A spirit offers a riddle; solve for a Clue.
    \item The Ninth Hour – sunset lasts twice as long. The GM gains 1 SB to spend in the next scene.
\end{enumerate}

\section{Factions of the Way}
\begin{tabu} to \linewidth {|X|c|c|c|X|}
\hline
\textbf{Faction} & \textbf{Influence} & \textbf{Stability} & \textbf{Crisis} & \textbf{Goal} \\
\hline
Fhara Caravan Courts & 5 & 6 & 4 & Keep water law intact \\
Sidhi Port Guilds & 5 & 4 & 6 & Stay free of Ashaan \\
Pereshi Highlander Lords & 6 & 5 & 5 & Maintain ancient trade charters \\
Kuvani Remount Leagues & 4 & 5 & 3 & Expand grazing rights \\
Tulkani Mother‑Roads & 3 & 4 & 2 & Keep the Way neutral \\
Dhaharan Viceroyalty & 6 & 5 & 5 & Balance Mountain and Admiralty \\
\hline
\end{tabu}

\clearpage
\chapter{Trade Along the Way}
\label{chap:trade}

Coin is heavy, easily counterfeited, and despised by the Tulkani. Along the Way of Silk, merchants trade in \textbf{Measures} – abstract units of value that represent a caravan’s cargo, reputation, and promises. A Measure is not a weight or a count; it is a \textit{negotiated agreement} on what a bolt of silk, a skin of water, or a safe passage is worth \textit{right now}. The system uses the same core currencies as \textit{Fate's Edge}: Boons, Strings, Debt Clocks, and Position.

\section{Core Concepts}

\begin{itemize}
    \item \textbf{Measures:} The basic unit of trade value. One Measure is roughly equivalent to a day’s labour, a meal for a caravan, or a small bribe. A caravan typically carries 1d6 + (Caravan Rating) Measures at the start of a journey.
    \item \textbf{Caravan Reputation (0–6):} A visible track that represents how trustworthy, skilled, or lucky a caravan is. Start at 2. Reputation modifies all trade rolls (add +1 die per point above 2, –1 die per point below).
    \item \textbf{Market Mood (d6):} Roll at each trading post or oasis. Affects prices (see table below).
    \item \textbf{Trade Goods:} Abstract categories: \textit{Food \& Water}, \textit{Silk \& Spice}, \textit{Iron \& Steel}, \textit{Luxury (perfume, glass, manuscripts)}, \textit{Slaves (illegal in most regions)}, \textit{Information (rumours, maps, letters)}.
\end{itemize}

\subsection*{Market Mood Table}
\begin{center}
\begin{tabular}{|c|l|}
\hline
\textbf{d6} & \textbf{Mood} \\
\hline
1 & Famine / War – Prices double, but danger increases (GM gains 2 SB). \\
2 & Scarcity – Prices +50\%, all trade rolls at –1 die. \\
3–4 & Normal – Prices as listed. \\
5 & Abundance – Prices –25\%, all trade rolls at +1 die. \\
6 & Festival / Truce – Prices –50\%, and the GM may offer a free String for a particularly good deal. \\
\hline
\end{tabular}
\end{center}

\subsection*{Trade Procedure}
\begin{enumerate}
    \item \textbf{Declare Intent:} The player states which goods they want to buy or sell, and how many Measures they are willing to risk.
    \item \textbf{Set Position:} Controlled if the market is normal, Dominant if they have a favourable String or a high Reputation, Desperate if the local faction hates them or if they are smuggling.
    \item \textbf{Roll:} \textbf{Presence + Sway} (haggling), \textbf{Wits + Streetwise} (finding a black market), or \textbf{Spirit + Lore} (negotiating with a cleric or judge). DV is set by the Mood (2–5).
    \item \textbf{Outcome:}
    \begin{itemize}
        \item \textbf{Clean Success:} Trade at listed price; gain +1 Reputation tick.
        \item \textbf{Success with SB:} Trade at price, but the GM gains 1 SB – the deal leaves a complication (a jealous rival, a forged seal, a delayed payment).
        \item \textbf{Partial:} Trade at –25\% (selling) or +25\% (buying); gain 1 Boon and a –1 Reputation tick.
        \item \textbf{Miss:} No trade; gain 2 Boons; the GM may start a Debt Clock [4] (e.g., “Owe the Factor”).
    \end{itemize}
\end{enumerate}

\section{Strings and Debts in Trade}

Merchants along the Way remember faces. A \textbf{String} can be spent to:
\begin{itemize}
    \item Gain +1 die on a trade roll.
    \item Demand a specific good that is not on the market (within reason).
    \item Convert a Miss into a Partial Success.
\end{itemize}

A \textbf{Debt Clock [4 or 6]} represents a loan, a broken promise, or a favour owed. When it fills, the creditor calls in the debt – often in the middle of a crisis. Clearing a segment requires an act of service (e.g., escort a caravan, deliver a message, sabotage a rival) or spending 2 Measures.

\section{Caravan Reputation – What It Does}
\begin{itemize}
    \item \textbf{Reputation 5–6:} “Trusted.” You may trade at one Position higher than normal. The first Miss in a scene becomes a Partial.
    \item \textbf{Reputation 3–4:} “Reliable.” No modifier.
    \item \textbf{Reputation 1–2:} “Suspicious.” The first Clean Success becomes a Success with SB (GM gains 1 SB).
    \item \textbf{Reputation 0:} “Untouchable.” No one will trade with you except at Desperate Position. You must perform a major service (a full session) to raise it back to 1.
\end{itemize}

\section{Special Trade Goods}

\subsection{Slaves (Illegal)}
Trading slaves is forbidden by the Covenant and the Temple of Light, but Ashaani and some Pereshi lords still deal in them. Selling a slave gives +2 Measures immediately but starts a \textbf{Witness Clock [6]}. When the clock fills, a Covenant judge or a Breath‑less agent confronts the party. Buying a slave and freeing them gives a permanent \textit{Freedman’s Seal} String (usable in Sidhi ports) but loses 1 Reputation in Ashaani circles.

\subsection{Information}
A rumour, a map, or a secret can be traded as a good. Information has no fixed price – the GM sets a Measure cost based on its value. Spending a String to “know a guy” can also be traded as Information.

\subsection{Water}
In the desert, water is more valuable than silk. A single Measure of water can be traded for two Measures of any other good during a drought. The GM may use the \textit{Dry Well} hazard to force water prices up.

\section{Example Trade Scene}

The party reaches a Fhara oasis. They have Caravan Reputation 3, and they want to sell three Measures of silk to buy water. The Market Mood is 4 (Normal). The player (Presence 3 + Sway 2 = 5 dice) decides to haggle.

\begin{enumerate}
    \item GM sets DV 3 (Normal), Position Controlled.
    \item Player rolls 5d10: 6, 7, 4, 2, 1 → 2 successes (6,7), one 1 → Success with SB.
    \item Outcome: They sell the silk at the listed price (3 Measures of silk = 3 Measures of water). The GM gains 1 SB.
    \item GM spends the SB: “The Fhara water‑judge notices that your silk has a slight stain – a moth hole. He demands you throw in a favour: deliver a sealed letter to the next oasis. The letter is a bill of debt.”
    \item The party now has a Debt Clock [4] “Carry the Letter”. If they fail, the water price triples at the next oasis.
\end{enumerate}

\section{Optional: Caravan Ledger (Mini‑game)}

For GMs who want more depth, each caravan can have a \textbf{Ledger} – a set of three Clocks:
\begin{itemize}
    \item \textbf{Supplies [6]:} Food, water, fodder. Advances when crossing difficult terrain, after a Miss on a trade roll, or when the GM spends 2 SB. When full, the caravan must buy supplies at double price or suffer Fatigue.
    \item \textbf{Debts [4]:} Obligations to local factions. Advances when the party takes a loan or breaks a promise. When full, a creditor appears – the party must pay 2 Measures or perform a service.
    \item \textbf{Opportunities [4]:} A rumour of a lucrative deal. Advances when the party spends a Boon to gather information, or when the GM wants to offer a hook. When full, the party can attempt a special trade: risk 1 Measure to gain 2 Measures on a successful roll, but lose it on a failure.
\end{itemize}

Each downtime (or after each leg), the party can “settle the ledger”: roll Wits + Lore (DV 3) to clear one segment from any one clock, or spend 1 Boon to clear two segments.

\section{Trade Goods Price List (in Measures)}
\begin{center}
\begin{tabular}{|l|c|}
\hline
\textbf{Good} & \textbf{Base Price} \\
\hline
1 day’s rations (per person) & 1 Measure (feeds 10) \\
\hline
Skin of water (per person per day) & 1 Measure \\
\hline
Bolt of silk (low quality) & 2 Measures \\
\hline
Bolt of silk (high quality) & 4 Measures \\
\hline
Spices (saffron, pepper, cinnamon – per pound) & 3 Measures \\
\hline
Iron ingot (10 lb) & 1 Measure \\
\hline
Steel sword & 4 Measures \\
\hline
Camel or horse & 6 Measures \\
\hline
Slave (unskilled) & 4 Measures (illegal) \\
\hline
Slave (skilled) & 8 Measures (illegal) \\
\hline
Map of a pass & 2 Measures \\
\hline
Rumour (verified) & 1 Measure \\
\hline
Covenant seal (forged) & 3 Measures \\
\hline
\end{tabular}
\end{center}

Prices are halved when selling, doubled when buying, and modified by Market Mood.

\section{Final Advice}
Trade in \textit{Fate's Edge} is not an economy simulation – it is a story engine. Use Measures, Strings, and Debt Clocks to create pressure and choices, not to track every copper. When the players cross the desert, they should worry about whether to bribe the water‑judge or save their Measures for the pass toll. And when they finally reach the brass gates of Dhahara, they should feel that every Measure they earned was bought with a promise, a risk, or a scar.

\begin{lorebox}
“The road is a ledger. Every oasis is a credit, every broken wheel a debt. The Tulkani tie knots to remember who owes whom. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. And the Sidhi – the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.”

--- \textit{Dusana of the Raven Road}
\end{lorebox}
\section{Hyaar — ``The Crossroads of Thrones''}
\label{sec:hyaar}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Hyaar is a land of petty kingdoms and fractured loyalties, a mountainous peninsula where the Amaranthine Sea meets the eastern trade winds. Its coasts are lined with Thepyrgosi colonies, Ashaani trading posts, and Sidhi lighthouse cities; its interior valleys harbour Pereshi highland lords, Fhara oasis‑clans, and native hill tribes who have never bowed to any empire. Every pass is a border, every bridge a treaty, and every inn a court of whispers. To travel Hyaar is to navigate a mosaic of jurisdictions where yesterday’s ally is today’s tax collector.

\vspace{2mm}
\textbf{What you notice first:} The smell of pine resin and sea salt. The way every fortress flies two banners – one for the lord who holds it, one for the empire that claims it. The constant lowing of goats on terraced hillsides.

\textbf{The one rule that kills newcomers:} Never carry a letter sealed with a single colour of wax. The colour tells which lord’s justice you trust. The wise use a ribbon of three.
\end{tcolorbox}

\subsection*{Prominent NPCs of Hyaar}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Arkadios & Thepyrgos Exarch & Keep the synod’s tithe flowing & He wears a key-shaped earring; it opens a wine cellar, not a door. \\
\hline
Lady Tamar & Pereshi highland lord & Protect her valley from lowland tax men & She keeps a library of ledgers; each is a different clan’s debt. \\
\hline
Emissary Davud & Fhara water‑clerk & Negotiate well‑rights for his oasis & He never drinks water unless it is poured by another. \\
\hline
Captain Heliodora & Sidhi lighthouse commander & Stop Ashaani slave raids & Her lantern code includes a signal for “false flag”. \\
\hline
Sheikh Faisal & Ashaani horse‑broker & Buy Hyaari warhorses for the Veiled Sisters & He has never been seen without his left hand gloved. \\
\hline
\end{longtable}

\subsection*{Names of Hyaar}
A polyglot blend: Thepyrgosi (-os, -ia), Ashaani (al-, -i), Pereshi (-stan, -abad), Fhara (bin, bint, al-), Sidhi (el-, -on), and native Hyaari (-oglu, -ova).

\section*{Spades — Places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Three‑Flag Inn} — At a crossroads; each flag guarantees safe passage from a different lord. \textit{The fourth flag is a shroud; it means the innkeeper witnessed a murder.}
\item \textbf{Abandoned Thepyrgos Watchtower} — Now a smuggler’s cache; the beacon still works. \textit{Light it, and the wrong people will answer.}
\item \textbf{Aqueduct Gap} — Broken bridge; locals use a rope ladder. \textit{The rope is knotted with the names of those who fell.}
\item \textbf{Pereshi Garden Shrine} — Tended by a silent hermit. \textit{The flowers bloom only at night; they smell of incense and iron.}
\item \textbf{Fhara Well Court} — Deep stone basin; a judge sits on a carpet. \textit{The judge’s scales are two earthenware cups. He fills them with water, not coins.}
\item \textbf{Sidhi Lighthouse} — Brass mirrors and a keeper who never blinks. \textit{The light is said to show the face of whoever last drowned nearby.}
\item \textbf{Ashaani Horse Market} — Fenced field outside a mud‑brick fort. \textit{The horses have no names – only brands. The brands are the names of their debtors.}
\item \textbf{Native Hill Fort} — Timber palisade; the chief’s hall is a repurposed temple. \textit{The temple’s god was a bear. The chief wears its skull at councils.}
\item \textbf{Coin‑House of the Three Seas} — A bank that honours letters from Ecktoria, Kahfagia, and Ashaan. \textit{The vault is underwater; the teller holds his breath to open it.}
\item[J] \textbf{Leper’s Causeway} — A stone path to a quarantined village. \textit{The villagers are not lepers – they are debtors hiding from three kingdoms.}
\item[Q] \textbf{Parliament Oak} — A hollow tree where the hill tribes meet. \textit{The tree’s leaves are always green, even in winter. The sap is warm.}
\item[K] \textbf{The Silver Ford} — A river crossing that changes sides each season. \textit{The toll‑keeper is a ghost. He accepts only promises.}
\item[A] \textbf{The Caravanserai of the Ninth Lamp} — A ruin where no lamp stays lit. \textit{The ninth lamp, when lit, opens a door to yesterday.}
\end{enumerate}

\section*{Hearts — People \& Factions}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Coin‑House Factor} — Wears a vest with twelve pockets; each holds a different currency. \textit{He knows the exchange rate of secrets.}
\item \textbf{Toll‑Keeper’s Widow} — Collects the bridge fee with a stone face. \textit{She pays the ghosts of her husband’s debtors with a small portion of every coin.}
\item \textbf{Thepyrgos Exarch} — Robed, bearded, always writing letters. \textit{His letters are in code; the code is a poem. The poem changes weekly.}
\item \textbf{Pereshi Highlander} — Carries a rifle and a book of verse. \textit{He shoots the first round, then recites a couplet. The couplet is the bullet’s name.}
\item \textbf{Fhara Water‑Judge} — Holds a brass dipper instead of a gavel. \textit{He decides disputes by tasting the water. Salt means guilt.}
\item \textbf{Sidhi Lighthouse Keeper} — Blind in one eye; the other sees only green. \textit{He claims the sea speaks in colours. Green means “treachery”.}
\item \textbf{Ashaani Slave‑Courser} — Dressed in silk, followed by two eunuchs. \textit{He never looks at the people he buys. He looks at their shadows.}
\item \textbf{Native Hill Chief} — Scarred, grey‑haired, missing three fingers. \textit{Each missing finger is a treaty he broke. He keeps the rings on a cord around his neck.}
\item \textbf{Kuvani Remount Master} — Speaks little, watches the horses. \textit{He can tell a lie by how a horse flicks its tail.}
\item[J] \textbf{Breath‑less Agent} — Carries no metal; wears wool and a wooden ring. \textit{The ring is hollow. It contains a single word – “witness”.}
\item[Q] \textbf{The Pilgrim of the Three Faiths} — Visits Everflame, Temple, and Covenant shrines in one day. \textit{He prays in silence. Each shrine thinks he belongs to the other.}
\item[K] \textbf{The Exiled Admiral} — Runs a shipyard in a hidden cove. \textit{Her flagship is a fishing boat; the cannons are hidden in the cargo holds.}
\item[A] \textbf{The Shepherd of the Pass} — A hooded figure who guides travellers through the mountains. \textit{He never takes payment – only a story. He never laughs at the stories.}
\end{enumerate}

\section*{Clubs — Complications/Threats}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Flag Dispute} — Two lords claim the same inn; you are the witness. \textit{The innkeeper has two bills; both are stamped with the same seal. The seal is a forgery.}
\item \textbf{Dry Well Court} — The water‑judge demands you fill the basin. \textit{The basin is a trial; you must bring water from a spring that disappeared a year ago.}
\item \textbf{Lighthouse out of Phase} — The Sidhi keeper signals a warning; your ship is in a reef. \textit{The warning was for the ship behind you. You are now the decoy.}
\item \textbf{Horse Market Brawl} — An Ashaani buyer refuses to pay. \textit{The horse is a stolen war‑horse; its brand matches a dead lord’s crest.}
\item \textbf{Toll Increase} — The widow raises the fee; she is collecting for a funeral. \textit{The funeral is yours if you do not pay.}
\item \textbf{Tax Audit} — The Coin‑House demands to see your ledgers. \textit{Your ledgers are written in lemon juice; heat reveals a second column – bribes.}
\item \textbf{Ghost Caravan} — A train of skeletal mules blocks the road at midnight. \textit{The mules’ cargo is sealed jars; the jars whisper “open us”.}
\item \textbf{Audience of the Hill Chief} — He demands a gift. \textit{The gift must be a secret about another tribe. Refuse, and you are the secret.}
\item \textbf{Slave Raid} — Ashaani slavers attack the village. \textit{The villagers are armed; they have been expecting you. You are the bait.}
\item[J] \textbf{The Ninth Lamp} — A stranger offers to light it. \textit{If you let them, yesterday’s dead will walk for an hour. They are not friendly.}
\item[Q] \textbf{Pereshi Blood Feud} — Two highland clans blame you for a stolen herd. \textit{The herd is three goats; the goats are the property of a third clan – the one that hired you.}
\item[K] \textbf{Thepyrgos Interdict} — The synod declares the pass cursed; all travel must stop. \textit{The interdict was bought by a rival caravan; the curse is a bucket of red paint.}
\item[A] \textbf{The Shepherd’s Price} — Your guide stops at a fork. \textit{“One road leads to safety, one to death. Tell me a story – and I will point.”}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Three‑Flag Pass} — A pennant that lowers tolls on all three lords’ roads. \textit{The colours fade in sunlight; the knots must be retied each dawn.}
\item \textbf{Water‑Judge’s Token} — A clay dipper; present it to skip a queue at any well. \textit{The dipper cracks if you lie about your need.}
\item \textbf{Lighthouse Charter} — The right to anchor in a Sidhi port without inspection. \textit{The charter is a brass key; it fits a lock on a sea‑chest, not a harbour gate.}
\item \textbf{Hill Chief’s Spear} — A ceremonial weapon; carrying it grants safe passage in the highlands. \textit{The spear is old; the blade is bronze. It cannot cut flesh – only oaths.}
\item \textbf{Coin‑House Credit} — A sealed letter; redeemable at any branch. \textit{The letter is written on both sides; the back is a map to a vault that does not exist.}
\item \textbf{Ashaani Warranty} — A bill of sale for a horse; the horse is not named. \textit{The warranty is a brand. Press it onto any horse, and the seller must honour it.}
\item \textbf{Pereshi Verse of Passage} — A poem; reciting it to a guard halves any toll. \textit{The poem has eight lines. The ninth is the toll‑keeper’s name.}
\item \textbf{Kuvani Remount Marker} — A horse‑hair token; exchange for a fresh mount at any station. \textit{The token is braided with a single white hair. That hair belonged to a horse that died on the road.}
\item \textbf{Breath‑less Cipher Ring} — A silver band; twist it to reveal a hidden message. \textit{The message is a name. The name is a debt. The debt is yours.}
\item[J] \textbf{The Exile’s Chart} — A map of hidden coves and smugglers’ stairs. \textit{The map is tattooed on the inside of a leather glove. Wear it, and the sea remembers you.}
\item[Q] \textbf{The Shepherd’s Compass} — A brass needle that points to the nearest safe hearth. \textit{The compass is broken; it points to where you were yesterday.}
\item[K] \textbf{The Caravanserai’s Ledger} — A log of every guest for a century. \textit{The margins contain comments. Your grandfather is mentioned – as a thief.}
\item[A] \textbf{The Ninth Lamp Shard} — A piece of the broken lamp; once per campaign, light it to ask a dead person one question. \textit{The dead person must be named in the ledger.}
\end{enumerate}

\section*{Quick use notes}
\begin{itemize}
\item Draw until all four suits appear. Highest rank sets main Clock (2–5:4, 6–10:6, J/Q/K:8, A:10).
\item Diamonds are narrative permits (passes, charters, tokens) that shift Position rather than call for a roll.
\item If any A appears, echo \textbf{three flags and nine lamps} – thresholds multiply, laws contradict, and the dead answer.
\end{itemize}

\subsection*{Hyaar Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Ford:} Toss a crust into the Silver Ford before crossing. Ignore the first \textit{Dry Well Court} penalty.
\item \textbf{Knot the Three Flags:} Tie the pennant’s cords together. Cancel \textit{Flag Dispute} or take –1 die on your next negotiation (your choice).
\item \textbf{Name to the Lighthouse:} Whisper a dead sailor’s name into the mirror. Gain +1 Effect when evading the Ashaani this scene, but mark Obligation to the Sidhi.
\end{itemize}

% ------------------------------------------------

\section{Galanina — ``The Sealed Coast''}
\label{sec:galanina}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Galanina is the ancestral homeland of the Sidhi, a rugged coast of limestone cliffs, hidden coves, and ancient ports. It is also the heart of the Covenant – the law‑faith of Mykkiel, whose judges and notaries have stamped their seal on every document for a thousand years. The Ashaani magocracy has tried to conquer Galanina twice; both times, the Covenant courts declared the invasion “procedurally void” and the armies withered from legal exhaustion. Today, Galanina is a quiet, suspicious land of lighthouse‑courts, archive‑fortresses, and families who keep their secrets in sealed jars.

\vspace{2mm}
\textbf{What you notice first:} The smell of old parchment and salt. The way every door has a brass seal – broken seals mean abandoned houses. The constant ticking of water‑clocks in public squares.

\textbf{The one rule that kills newcomers:} Never open a sealed jar that does not have a witness. The Covenant counts the opening as an “unwitnessed testimony” – admissible only against you.
\end{tcolorbox}

\subsection*{Prominent NPCs of Galanina}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Magistrate Elara & Covenant Judge of the Salt Circuit & Enforce maritime law & She carries a water‑clock; it runs backwards when she lies. \\
\hline
Keeper Solon & Archivist of the Unforgotten Ledger & Protect the original Covenant texts & He has never been seen eating; his inkwell is always full. \\
\hline
Captain Idris & Sidhi lighthouse commander & Prevent Ashaani slavers from landing & His left eye is a brass lens; he can read the stars even in fog. \\
\hline
Envoys Naveed & Breath‑less agent & Recruit exiles to resist the Veiled Sisters & He has three names; each is for a different jurisdiction. \\
\hline
\end{longtable}

\subsection*{Names of Galanina}
Sidhi and Covenant names: Elara, Solon, Idris, Naveed, Tamar, Silas, Malachi, Jael.

\section*{Spades — Places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{The Archive of Sealed Doors} — A fortress filled with ledgers; each door is locked with a different judge’s seal. \textit{The seals are not wax – they are brass rings. Each ring is a different size.}
\item \textbf{Lighthouse of the Unfaithful Dead} — A tower where the lantern burns green. \textit{The light shows the faces of those who died in Ashaani chains. They do not blink.}
\item \textbf{Covenant Court of the Salt Marsh} — A wooden platform on stilts. \textit{The tide rises during trials; if the water touches the judge’s feet, the verdict is “drowned”.}
\item \textbf{Sidhi Hidden Cove} — A beach accessible only through a sea‑cave. \textit{The cave walls are covered in tally marks – each mark is a slave freed.}
\item \textbf{The Unforgotten Ledger’s Vault} — A subterranean room; the walls are shelves of sealed jars. \textit{Each jar contains a broken oath. The jars whisper if you listen.}
\item \textbf{The Bitter Well} — A freshwater spring that tastes of iron. \textit{It is said that the first Covenant judge drowned a liar there. The water still remembers his name.}
\item \textbf{The Waxworks of the Anathema} — A building where excommunicated judges are immortalised in wax. \textit{The wax figures move at night. They write their own appeals on the walls.}
\item \textbf{The Admirals’ Gallery} — A portico with statues of Sidhi ship captains. \textit{Each statue holds a stone lantern; the lanterns glow when a descendant is at sea.}
\item \textbf{The Court of Broken Chains} — An open‑air ruin; the floor is a mosaic of shattered manacles. \textit{The mosaic is a map; the missing pieces are the locations of secret Anchor sites.}
\item[J] \textbf{The Silent Scriptorium} — A monastery where scribes copy old judgments. \textit{They write with their left hands; their right hands are for swearing oaths.}
\item[Q] \textbf{The Harbour of the Twice‑Drowned} — A port built over a sunken city. \textit{At low tide, you can see the rooftops. The bells of the drowned church still ring.}
\item[K] \textbf{The Judge’s Pool} — A reflecting basin in a courtyard. \textit{The water shows not your face but the face of the person you have wronged.}
\item[A] \textbf{The Archive of the Ninth Seal} — A locked room that no judge has entered. \textit{The seal is a single word: “Witness”. The door is warm.}
\end{enumerate}

\section*{Hearts — People \& Factions}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Covenant Magistrate} — Robed, bald, always carrying a scale. \textit{The scale’s pans are empty; she weighs arguments, not evidence.}
\item \textbf{Sidhi Lighthouse Keeper} — Blind in one eye; the other sees only the sea. \textit{He lights the lamp with a piece of the sun – or so he claims.}
\item \textbf{Archive Scribe} — Ink‑stained fingers, silent as a fish. \textit{She has transcribed every death sentence for thirty years. She has never repeated a name.}
\item \textbf{Breath‑less Agent} — Dressed as a fisher, but his hands are uncalloused. \textit{He speaks in a whisper; his words appear on paper without ink.}
\item \textbf{Shipwright of the Freedmen} — Builds boats for escaped slaves. \textit{Each boat is named after a dead family member. The names are never painted – only carved.}
\item \textbf{Salt Farmer} — Harvests salt from pans; her hands are cracked. \textit{She pays her taxes in salt; the Covenant accepts it as “liquid testimony”.}
\item \textbf{Debt Collector of the Anvil} — A dwarf from Aeler, authorised to collect on maritime loans. \textit{He has a hammer; he knocks once for a warning, twice for a summons, thrice for a contract.}
\item \textbf{Pereshi Exile} — A scholar who fled a blood feud. \textit{He writes poetry on the walls of the Archive. The archivists have not erased it.}
\item \textbf{Fhara Water‑Broker} — Buys well‑rights and resells them. \textit{He never drinks water. Only wine. He says water is “unsafe”.}
\item[J] \textbf{Ashaani Defector} — A former slave‑courser who now works for the Breath‑less. \textit{He wears a mask to hide his face. The mask is a slave’s registration seal.}
\item[Q] \textbf{The Anchorite of the Ninth Seal} — A hermit who guards the sealed door. \textit{He has no shadow. The door’s warmth keeps him alive.}
\item[K] \textbf{The Admiral of the Salt Fleet} — Commands a flotilla of fishing boats. \textit{The boats are armed with Covenant‑blessed ballistae. They hunt slavers.}
\item[A] \textbf{The Voice of Mykkiel} — An oracle who speaks only in citations. \textit{Her eyes are silver; they reflect the text of the Covenant.}
\end{enumerate}

\section*{Clubs — Complications/Threats}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Seal Breach} – A jar is found open; the oath inside is now “witnessed”. \textit{The oath was a promise of safe passage. It is now void.}
\item \textbf{Tide Court} – The water rises during a trial; the judge declares “drowned” as the only verdict. \textit{The accused is not present – they fled. You are the witness.}
\item \textbf{Ashaani Raid} – Slave ships are sighted. \textit{The Sidhi lighthouse keeper asks you to hold the light. Hold it too long, and they will see you.}
\item \textbf{Covenant Audit} – A magistrate demands to see your contracts. \textit{Your contract with the Fhara was never witnessed. It is void. You are a debtor.}
\item \textbf{Forged Seal} – A document you carry is discovered to be a forgery. \textit{The seal is yours – or a perfect copy. You are the forger.}
\item \textbf{Archive Fire} – A lamp falls; ledgers burn. \textit{The smoke forms names. Those names are the ones the Covenant wants erased.}
\item \textbf{Breath‑less Betrayal} – An agent sells you to the Sisters. \textit{The agent is a slave in the Sisters’ household. The betrayal was a test.}
\item \textbf{The Ninth Seal Opens} – The sealed door cracks. \textit{From inside comes a voice: “Who witnesses?” You must answer – or be sealed.}
\item \textbf{Salt Flood} – A storm breaks the pans; the town is buried in salt. \textit{The salt is pure. It is also a tax record. You owe.}
\item[J] \textbf{Drowned Verdict} – A corpse is found with a Covenant seal on its tongue. \textit{The verdict was “exile”. The sea delivered it.}
\item[Q] \textbf{The Silent Scriptorium’s Curse} – Scribes begin writing your name in the margins. \textit{The margins are for corrections. Your name is now a correction.}
\item[K] \textbf{The Admiral’s Debt} – She asks you to sink a rival’s boat. \textit{The rival is a Covenant judge. The boat is a courthouse.}
\item[A] \textbf{The Witness’s Demand} – Mykkiel’s oracle asks: “What did you bury?” \textit{You must answer. The answer will be sealed in a jar.}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Covenant Seal Ring} – Press it into wax to make any document “witnessed”. \textit{The ring is warm. It cools if you lie.}
\item \textbf{Lighthouse Charter} – Right to anchor in any Sidhi port. \textit{The charter is a key; it fits a lock in the lighthouse keeper’s chest.}
\item \textbf{Unforgotten Ledger Page} – A single leaf; it contains a name and a debt. \textit{The debt is transferable. Sign it over to avoid payment.}
\item \textbf{Freedman’s Seal} – Proof of emancipation; Ashaani slavers cannot touch you. \textit{The seal is a brand. It glows when a slave ship is near.}
\item \textbf{Archive Key} – Opens one door in the sealed Archive. \textit{The key is a finger bone. The Archive keeps the corresponding finger.}
\item \textbf{Salt Farmer’s Tithe} – A bag of salt; legal tender in any Covenant court. \textit{The salt is bitter. It tastes of tears.}
\item \textbf{Breath‑less Cipher} – A codebook; allows you to read secret messages. \textit{The codebook is written in disappearing ink. It reappears when wet.}
\item \textbf{Anchorite’s Whisper} – A single word from the Ninth Seal. \textit{The word is a name. Speak it to open any lock – once.}
\item \textbf{Admiral’s Favour} – A fishing boat with a hidden ballista. \textit{The ballista fires salt‑dipped bolts. They are legal evidence.}
\item[J] \textbf{The Voice’s Citation} – A legal precedent; quoting it wins one argument. \textit{The citation is a number. The number is your age.}
\item[Q] \textbf{The Shipwright’s Boat} – A vessel built for escape. \textit{The boat has no name. It cannot be tracked by magic.}
\item[K] \textbf{The Salt Fleet’s Flag} – Fly it to be immune from Ashaani attack. \textit{The flag is a shroud. It was worn by an Admiral’s daughter.}
\item[A] \textbf{The Ninth Seal Fragment} – A piece of the broken door. \textit{Once per campaign, touch it to a sealed jar to learn the oath inside.}
\end{enumerate}

\section*{Quick use notes}
\begin{itemize}
\item Draw until all four suits appear. Highest rank sets main Clock.
\item Diamonds are legal instruments (seals, charters, keys) that shift Position.
\item If any A appears, echo \textbf{seals and salt} – jars crack, tides rise, and the Covenant demands a witness.
\end{itemize}

\subsection*{Galanina Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Salt:} Toss a pinch into a pan. Ignore first \textit{Seal Breach} penalty.
\item \textbf{Knot the Ledger:} Tie a ribbon around a page. Cancel \textit{Archive Fire} or take –1 die on your next legal roll.
\item \textbf{Name to the Well:} Whisper a dead judge’s name into the Bitter Well. Gain +1 Effect when arguing a case this scene, but mark Obligation to the Covenant.
\end{itemize}

% ------------------------------------------------

\section{Pereshi Highlands — ``The Roof of the Way''}
\label{sec:pereshi}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Pereshi highlands are a vast, windswept plateau, cut by deep river valleys and crowned with snow‑capped peaks. Ancient fire temples stand beside caravanserais; terraced vineyards climb the hillsides. The people are scholars, shepherds, and traders – proud, suspicious, and bound by a code of hospitality that demands three days of feasting for any guest, and three generations of vengeance for any insult. The Pereshi have no king; their land is a patchwork of petty lords, each ruling from a stone fortress, each claiming descent from the mythic Sun Kings of the First Age.

\vspace{2mm}
\textbf{What you notice first:} The smell of woodsmoke and roasting lamb. The way every valley has a different accent. The constant call to prayer from fire‑temple towers – but the prayers are in a language no one speaks outside the highlands.

\textbf{The one rule that kills newcomers:} Never refuse a second cup of tea. The first is hospitality; the second is a test of honour. The third is a promise.
\end{tcolorbox}

\subsection*{Prominent NPCs of the Pereshi Highlands}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Lord Rostam & Highland petty king & Unite the valleys against lowland tax men & He keeps a cage of doves; each dove carries a different treaty. \\
\hline
Flame‑Keeper Zarifa & Fire‑temple priestess & Preserve the old rites & She has never been seen without a rose – fresh or dried. \\
\hline
Tarkhan & Caravan master & Secure the passes & He has three camels; each is named after a dead wife. \\
\hline
Mirza & Kuvani remount broker & Sell horses to the Pereshi & He speaks Pereshi with a Ykrul accent; the Pereshi find it charming. \\
\hline
\end{longtable}

\section*{Spades — Places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Fire Temple of the Unquenched} – A stone tower with a copper dome. \textit{The fire inside is said to be the first flame. It has never gone out – but it flickers when a king dies.}
\item \textbf{The Caravanserai of the Three Valleys} – A walled courtyard; three gates, each from a different valley. \textit{The gate you enter determines the price you pay.}
\item \textbf{Vineyard of the Blind Vintner} – Terraced hillside; the vintner lost his sight to a curse. \textit{He tastes the grapes to know the weather. He is never wrong.}
\item \textbf{The Library of Unfinished Books} – A scriptorium where scribes copy texts that were burned. \textit{The books are blank. The scribes are ghosts.}
\item \textbf{The Bridge of the Ninth Verse} – A stone arch over a chasm. \textit{Each stone is carved with a line of poetry. The ninth verse is missing – it was stolen by a king.}
\item \textbf{The Shepherds’ Moot} – A circle of standing stones on a hilltop. \textit{The shepherds gather here to settle disputes. The stones have ears.}
\item \textbf{The Salt Caravan’s Grave} – A dry lake bed; the bones of a thousand camels. \textit{The salt is still white. It tastes of old blood.}
\item \textbf{The Hospice of the Lost Steps} – A monastery where pilgrims who lost their way are healed. \textit{The healers ask no payment – only a story of how you got lost.}
\item \textbf{The Sun King’s Throne} – A carved stone seat on a peak. \textit{Whoever sits there at dawn can see three valleys. The curse is that they cannot leave.}
\item[J] \textbf{The Forge of the Silent Smith} – A cave where a hermit makes swords. \textit{The swords have no edge. They are legal documents, folded steel.}
\item[Q] \textbf{The Well of the Ancestors} – A deep pool in a grove of cypress trees. \textit{The water is black. Drink it, and you remember a past life – but you forget one year of this one.}
\item[K] \textbf{The Fire Temple of the Ninth Coal} – A sanctuary where the flame is kept in a brass brazier. \textit{The brazier has nine coals. The ninth is never lit.}
\item[A] \textbf{The Pass of the Unforgiven} – A narrow defile where a treaty was broken. \textit{The wind carries the names of the oath‑breakers. It never stops.}
\end{enumerate}

\section*{Hearts — People \& Factions}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Flame‑Keeper} – Dressed in white, carries a brass dipper. \textit{He pours oil into the fire; each drop is a prayer. He prays for the dead.}
\item \textbf{Vineyard Owner} – Hands stained purple; a gold earring. \textit{She can tell the age of a wine by the number of lies told during its fermentation.}
\item \textbf{Caravan Master} – Worn leather, soft voice. \textit{He never carries a weapon. His camels are his guards.}
\item \textbf{Shepherd King} – Young, wearing a cloak of uncut wool. \textit{He counts his sheep by naming them. He has never lost a name.}
\item \textbf{Kuvani Horse Trader} – Braided hair, silver earrings. \textit{He names his horses after rivers. The rivers are the ones he has crossed.}
\item \textbf{Pereshi Scholar} – Ink‑stained fingers, half‑moon spectacles. \textit{He carries a book of blank pages. He writes only in his memory.}
\item \textbf{Temple Beggar} – Blind, always sitting by the fire. \textit{He can see the future in the flames. He charges a single coal.}
\item \textbf{Oath‑Breaker’s Ghost} – Visible only at twilight. \textit{He repeats the words of the broken treaty. He has been repeating them for a century.}
\item \textbf{Fhara Water‑Seller} – A stranger in the highlands, selling water from the lowlands. \textit{His water is expensive. It is also the only water that tastes of home.}
\item[J] \textbf{The Poet of the Ninth Verse} – An old man who claims to have stolen the missing verse. \textit{He recites it for a cup of tea. The tea must be made with snowmelt.}
\item[Q] \textbf{The Healer of the Lost Steps} – A woman with a staff of olive wood. \textit{She never asks your name. She knows it anyway.}
\item[K] \textbf{The Silent Smith} – A giant of a man, deaf and mute. \textit{He communicates by striking his anvil in code. The code is a forgotten language.}
\item[A] \textbf{The Sun King’s Shadow} – A figure that sits on the throne at night. \textit{It is not the king. It is the memory of the king. It gives bad advice.}
\end{enumerate}

\section*{Clubs — Complications/Threats}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Treaty Broken} – A lord declares war over a stolen goat. \textit{The goat was a witness. The treaty’s ninth clause names goats as “non‑hostile”.}
\item \textbf{Fire Flickers} – The temple flame gutters; a sign of a coming death. \textit{The death is yours – unless you pay a coal to the beggar.}
\item \textbf{Landslide} – The pass is closed; you must take the Unforgiven Pass. \textit{The wind there has names. Yours is among them.}
\item \textbf{Poet’s Demand} – The old man asks for a new verse. \textit{The verse must be true. If it is not, he will not speak again.}
\item \textbf{Horse Sickness} – A disease spreads through the Kuvani remounts. \textit{The horses dream of water. They wake thirsty for blood.}
\item \textbf{Blood Feud} – Two clans blame you for a shepherd’s death. \textit{The shepherd died of old age. The clans do not care.}
\item \textbf{Sheep Rustlers} – Your flock is stolen. \textit{The rustlers are children. They are hungry. Feeding them costs more than the sheep.}
\item \textbf{Drought} – The wells are low; the Fhara water‑seller raises his price. \textit{His price is a promise. The promise is to forget something.}
\item \textbf{Ancestor’s Demand} – A ghost appears in the well. \textit{It asks for a name. The name must be a dead enemy’s.}
\item[J] \textbf{The Ninth Coal Lit} – Someone lights the forbidden coal. \textit{The fire turns blue. The blue flame shows the future – but the future is always a funeral.}
\item[Q] \textbf{The Unforgiven Pass’s Toll} – The wind demands a name. \textit{Give a name, and the wind will carry it to the oath‑breakers. They will come.}
\item[K] \textbf{The Silent Smith’s Price} – He offers a sword. \textit{The price is a memory. He takes it with his tongs.}
\item[A] \textbf{The Sun King’s Curse} – Someone sits on the throne. \textit{They cannot leave. The party must find a new king – or break the throne.}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Flame‑Keeper’s Coal} – A single coal; light it to see one minute into the future. \textit{The coal is warm. It will not burn your hand.}
\item \textbf{Vintner’s Cask} – A bottle of wine; drink it to forget a failure. \textit{The failure is erased from your memory. The wine tastes of tears.}
\item \textbf{Caravan Master’s Map} – A chart of secret passes. \textit{The map is written on a camel’s hide. The hide is from a camel that died of thirst.}
\item \textbf{Shepherd King’s Staff} – A wooden rod; present it to any shepherd for safe passage. \textit{The staff is carved with the names of the sheep.}
\item \textbf{Kuvani Remount Token} – A horsehair bracelet; exchange for a fresh horse. \textit{The bracelet is braided with a promise. Break the promise, and the horse will throw you.}
\item \textbf{Pereshi Scholar’s Book} – A blank book; write a truth in it, and it becomes law for one day. \textit{The truth must be witnessed. The book is the witness.}
\item \textbf{Temple Beggar’s Vision} – A prophecy; once per session, the GM must answer one yes/no question about the future. \textit{The answer is always true – but always incomplete.}
\item \textbf{Oath‑Breaker’s Name} – A single name; speak it to void one contract. \textit{The contract is voided for both parties. The name is the price.}
\item \textbf{Poet’s Ninth Verse} – The missing verse; recite it to calm any spirit. \textit{The verse has no rhyme. The spirits prefer it that way.}
\item[J] \textbf{Healer’s Staff} – Touch it to remove one Condition. \textit{The Condition is transferred to the staff. The staff will break after three uses.}
\item[Q] \textbf{Silent Smith’s Sword} – A legal document folded into steel; present it to win any legal duel. \textit{The sword has no edge. It cuts oaths, not flesh.}
\item[K] \textbf{The Sun King’s Crown} – A circlet of bronze; wear it to command one valley. \textit{The crown is cold. It warms only when you speak truth.}
\item[A] \textbf{The Ninth Coal’s Ashes} – A pinch of ash; blow it into a fire to ask one question of the dead. \textit{The dead answer in smoke. The smoke writes on the wall.}
\end{enumerate}

\section*{Quick use notes}
\begin{itemize}
\item Draw until all four suits appear. Highest rank sets main Clock.
\item Diamonds are physical tokens (coals, staffs, ashes) that grant narrative permissions.
\item If any A appears, echo \textbf{fire and stone} – flames that judge, passes that remember, and kings who cannot leave.
\end{itemize}

\subsection*{Pereshi Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Fire:} Crumble a crust into the flame. Ignore first \textit{Fire Flickers} penalty.
\item \textbf{Knot the Verse:} Tie a ribbon around the Poet’s page. Cancel \textit{Treaty Broken} or take –1 die on your next oath.
\item \textbf{Name to the Well:} Whisper an ancestor’s name into the black water. Gain +1 Effect when fleeing a blood feud, but mark Obligation to the Flame‑Keeper.
\end{itemize}

% ------------------------------------------------

\section{Fhara — ``The Oasis Courts''}
\label{sec:fhara}

(Refer to the Fhara entry in the Way of Silk expansion for full generator; a condensed version is provided here for completeness. See \textit{Way of Silk}, Chapter 3.)

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Fhara dwell on a great, sun‑baked peninsula where scattered oases and dry riverbeds connect the coast to the interior. Their cities are built around wells and date groves, and their caravans are the arteries of the spice trade. Water is law; the well‑keepers are judges, and the courthouse is the shade of a palm. The Fhara are merchants, poets, and negotiators – they will bargain for an hour over a cup of coffee, and a day over a camel. Their honour is measured in generosity, and their vengeance is measured in litigation.

\vspace{2mm}
\textbf{What you notice first:} The smell of coffee and dates. The way every well has a stone bench for the judge. The silence when the sun passes directly overhead – no one moves, no one speaks.

\textbf{The one rule that kills newcomers:} Never refuse a third cup of coffee. The first is welcome, the second is negotiation, the third is a contract. Refuse it, and you have broken an oath.
\end{tcolorbox}

\subsection*{Prominent NPCs of the Fhara}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Sheikh Mansur & Oasis lord & Keep his well’s water rights & He has not tasted water in ten years – only coffee. \\
\hline
Fatima al‑Wadi & Water‑judge & Settle disputes without bloodshed & She carries a dipper made from a date‑palm leaf; it rots after each case. \\
\hline
Rashid & Caravan master & Find new routes & He has three wives; each is from a different oasis. The wives do not speak. \\
\hline
Layla & Coffee‑house poet & Sell verses for coin & Her poems are in code; the code is the price of coffee. \\
\hline
\end{longtable}

\section*{Spades — Places}
(Refer to Way of Silk; selected entries)
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Well Court} – A stone bench under a palm; a judge with a dipper. \textit{The dipper has a chip. The chip is from a sword that broke a well.}
\item \textbf{Coffee Bazaar} – Stalls under canvas; the smell of roasting beans. \textit{The best coffee is served in a chipped cup. The chip is a signature.}
\item \textbf{Date Grove} – Rows of palms; the ground is soft with fallen fruit. \textit{The fruit ferments at night. The drunk monkeys are the guards.}
\item \textbf{Caravanserai of the Three Waters} – A courtyard with three wells: sweet, brackish, and bitter. \textit{The bitter well is for prisoners. They drink once, then confess.}
\end{enumerate}

\section*{Hearts — People \& Factions}
\section*{Clubs — Complications/Threats}
\section*{Diamonds — Rewards/Leverage}
(Refer to Way of Silk; full table there.)

\subsection*{Fhara Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Well:} Crumble a crust at the stone bench. Ignore first \textit{Dry Well Court} penalty.
\item \textbf{Knot the Dipper:} Tie a date‑palm fibre around the handle. Cancel \textit{Coffee Contract} or take –1 die on your next negotiation.
\item \textbf{Name to the Grove:} Whisper a dead merchant’s name into the roots. Gain +1 Effect when haggling, but mark Obligation to the Oasis Lord.
\end{itemize}

% ------------------------------------------------

\section{Kuvani — ``The Remount Steppe''}
\label{sec:kuvani}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Kuvani range across the eastern steppe, a vast grassland east of the Valewood. They are a relatively new people in the region, their clan confederations only recently rising to prominence. They share the steppe with the Ykrul – but where the Ykrul are poets of geometry and oath‑strings, the Kuvani are pragmatists of the remount and the caravan. Their horsemanship and remount stations make them indispensable to any caravan that crosses the open plain. The Kuvani have no cities; their courts are circles of saddles, and their laws are sung.

\vspace{2mm}
\textbf{What you notice first:} The smell of horse sweat and dry grass. The way the horizon seems to move when you look away. The constant sound of bells on harnesses.

\textbf{The one rule that kills newcomers:} Never ride a Kuvani horse without asking its name. The horse will answer – by throwing you.
\end{tcolorbox}

\subsection*{Prominent NPCs of the Kuvani}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Alp Arslan & Remount master & Expand grazing rights & He has never been seen sleeping; his eyes are always on the horizon. \\
\hline
Borte & Wind‑singer & Keep the songs of the clans & She has a horse with a white blaze; the blaze is a map of the steppe. \\
\hline
Temur & Caravan guide & Find the fastest routes & He names his dogs after dead enemies. The dogs remember. \\
\hline
Seljuk & Ykrul envoy & Negotiate shared pastures & He carries a hostage string; the string has nine knots. \\
\hline
\end{longtable}

\section*{Spades — Places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Remount Station} – A felt tent with a corral; horses for hire. \textit{The horses have no names – only brands. The brands are the names of lost children.}
\item \textbf{Wind‑Singer’s Circle} – A ring of saddles; the singer sits in the centre. \textit{The songs are not sung – they are hummed. The hum is the wind.}
\item \textbf{Kurgan of the Broken Oath} – A burial mound with a stone horse. \textit{The stone horse is missing a leg. The leg is a treaty.}
\item \textbf{Grass Maze} – A section of the steppe where the grass grows in spirals. \textit{The maze is a map of the stars – but the stars have changed.}
\item \textbf{Salt Pan of the Ghost Herd} – A dry lake; the salt is crusted with hoofprints. \textit{The prints are from a herd that died a century ago. They appear every spring.}
\item \textbf{Felt Village} – A temporary camp of round tents. \textit{The tents face the wind; the wind is the clan’s direction.}
\item \textbf{Horse Cemetery} – A field of skulls, each on a wooden stake. \textit{The stakes are carved with names. The names are the horses’ riders.}
\item \textbf{Bridge of the Ninth Hardship} – A log over a ravine; the log is replaced each year. \textit{The old log is burned. The ash is the clan’s memory.}
\item \textbf{Stone of the Hostage String} – A standing stone with notches. \textit{Each notch is a hostage taken. The hostages are carved into the stone.}
\item[J] \textbf{The Iron Corral} – A pen made of salvaged sword blades. \textit{The blades are from a war that never ended. The horses are afraid of them.}
\item[Q] \textbf{The Sky‑Road Marker} – A cairn painted blue. \textit{The blue is from a stone that fell from the sky. The stone is still warm.}
\item[K] \textbf{The Khagan’s Felt Throne} – A tent within a tent; the seat is a saddle. \textit{The saddle has no stirrups. The Khagan does not need them.}
\item[A] \textbf{The Steppe’s Edge} – Where the grass ends and the desert begins. \textit{The line is a scar. The scar is a treaty.}
\end{enumerate}

\section*{Hearts — People \& Factions}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Remount Keeper} – Calloused hands, squinting eyes. \textit{He knows the price of a horse in salt, in silk, or in stories.}
\item \textbf{Wind‑Singer} – Grey‑haired, blind, but never lost. \textit{She hums the direction to the nearest water. The hum changes with the wind.}
\item \textbf{Horse Thief} – Young, fast, missing two fingers. \textit{The fingers were a bet. He lost the bet – but not the horses.}
\item \textbf{Ykrul Envoy} – Braided hair, a hostage string on his wrist. \textit{The string’s knots are for each season he has spent among the Kuvani.}
\item \textbf{Caravan Cook} – Fat, cheerful, always stirring a pot. \textit{He puts a bell in the pot; the bell rings when the meal is ready. The bell is a warning.}
\item \textbf{Grass Reader} – A hermit who lives in a felt hut. \textit{He reads the future in the way the grass bends. The future is always wind.}
\item \textbf{Fhara Merchant} – A stranger, selling dates and water. \textit{He pays his caravan guard in coffee. The guard is always awake.}
\item \textbf{Aeler Prospector} – A dwarf looking for iron. \textit{He carries a tuning fork; he taps it on stones. The stone that hums is ore.}
\item \textbf{Sidhi Runaway} – A young woman, fleeing Ashaan. \textit{She hides among the horses. The horses do not betray her.}
\item[J] \textbf{The Khagan’s Shadow} – A silent figure in black felt. \textit{He never speaks; he writes commands in the dust. The dust blows away.}
\item[Q] \textbf{The Horse‑Whisperer} – A child who can calm any stallion. \textit{The child speaks in neighs. The horses answer in riddles.}
\item[K] \textbf{The Ghost of the Broken Oath} – A rider with no shadow. \textit{He follows caravans that broke a treaty. He does not attack – only watches.}
\item[A] \textbf{The Steppe’s Voice} – An oracle who speaks when the wind dies. \textit{The oracle is a rock. The rock has a crack. The wind speaks through it.}
\end{enumerate}

\section*{Clubs — Complications/Threats}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Horse Sickness} – A cough spreads through the remounts. \textit{The cough is a curse. The curse was placed by a Ykrul bone‑singer.}
\item \textbf{Grass Fire} – The steppe burns; the smoke is black. \textit{The smoke spells a name. The name is the one who lit the fire.}
\item \textbf{Hostage String Cut} – A Ykrul envoy’s string is broken. \textit{The clan demands a new hostage. The party must choose.}
\item \textbf{Rasputitsa} – The mud swallows wagons; progress slows. \textit{The mud is warm. Something is buried underneath.}
\item \textbf{Stampede} – A herd of wild horses crashes through camp. \textit{The horses are not wild. They are fleeing something worse.}
\item \textbf{Wind‑Singer’s Silence} – She stops humming. \textit{The wind stops too. The silence is a warning.}
\item \textbf{Water Poisoned} – A well is found with dead animals nearby. \textit{The poison is salt. The salt is from a cursed pan.}
\item \textbf{Felt Village Gone} – A camp is empty; the tents are still standing. \textit{The felt is warm. The people vanished at noon.}
\item \textbf{Khagan’s Demand} – He asks for a tribute. \textit{The tribute is a story. The story must be true – or the Khagan will know.}
\item[J] \textbf{The Iron Corral’s Curse} – The sword‑blade fence begins to bleed. \textit{The blood is from the war. The war never ended.}
\item[Q] \textbf{The Sky‑Road Marker’s Fall} – The blue cairn collapses. \textit{The stone inside is a meteor. The meteor is hot. It whispers.}
\item[K] \textbf{The Ghost Herd’s Charge} – Spectral horses trample the camp. \textit{They are looking for their riders. The riders are dead.}
\item[A] \textbf{The Steppe’s Edge Recedes} – The grass grows into the desert. \textit{The desert is retreating. Something is driving it back.}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Remount Chain Mark} – A metal tag; exchange for a fresh horse. \textit{The tag is a horse’s tooth. The tooth is from a horse that died in battle.}
\item \textbf{Wind‑Singer’s Hum} – A recorded melody; hum it to navigate any storm. \textit{The hum is on a bone flute. The flute was carved from a rib.}
\item \textbf{Hostage String} – A braided cord; present it to demand a truce. \textit{The cord is hair. The hair is from a Khagan’s daughter.}
\item \textbf{Grazing Compact} – A written agreement; allows pasturing on clan land. \textit{The compact is on sheepskin. The sheep was eaten by the Khagan.}
\item \textbf{Felt Tent Pass} – A token; safe lodging in any camp. \textit{The token is a felt circle. The circle has a hole – for smoke.}
\item \textbf{Kurgan Key} – A stone tablet; opens the Horse Cemetery gate. \textit{The gate is a skull. The key fits the eye socket.}
\item \textbf{Iron Corral Nail} – One sword blade; use it to seal an oath. \textit{The blade is rusty. The rust is from broken promises.}
\item \textbf{Sky‑Road Stone} – A chip of blue stone; carry it to avoid getting lost. \textit{The stone is warm. It points to the north.}
\item \textbf{Khagan’s Whistle} – A brass whistle; summon a remount. \textit{The whistle is a horse’s phalange. The horse was the Khagan’s favourite.}
\item[J] \textbf{The Ghost Herd’s Bridle} – A leather strap; use it to command one spirit horse. \textit{The bridle is from a horse that drowned. The spirit is still wet.}
\item[Q] \textbf{The Wind‑Singer’s Staff} – A wooden rod; plant it to stop the wind. \textit{The staff is a tent pole. The tent was the singer’s home.}
\item[K] \textbf{The Khagan’s Saddle} – A felt pad; ride in it to gain +1 Position on all mounted rolls. \textit{The saddle has no stirrups. You must learn to ride without them.}
\item[A] \textbf{The Steppe’s Edge Stone} – A piece of the boundary; touch it to make the grass grow. \textit{The stone is a tooth. The tooth is from the Steppe itself.}
\end{enumerate}

\section*{Quick use notes}
\begin{itemize}
\item Draw until all four suits appear. Highest rank sets main Clock.
\item Diamonds are equestrian and territorial tokens (marks, strings, compacts) that shift Position.
\item If any A appears, echo \textbf{horses and wind} – remounts that remember, fires that spell names, and edges that move.
\end{itemize}

\subsection*{Kuvani Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Horses:} Crumble a crust into the corral. Ignore first \textit{Horse Sickness} penalty.
\item \textbf{Knot the Hostage String:} Tie a new knot. Cancel \textit{Hostage String Cut} or take –1 die on your next negotiation with Ykrul.
\item \textbf{Name to the Wind:} Whisper a dead horse’s name into the grass. Gain +1 Effect when riding this scene, but mark Obligation to the Wind‑Singer.
\end{itemize}

\clearpage
\chapter*{Colophon}
\addcontentsline{toc}{chapter}{Colophon}

\vspace{2\baselineskip}
\begin{lorebox}
The road is a ledger. Every oasis is a credit, every broken wheel a debt. The Tulkani tie knots to remember who owes whom. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. The Kuvani count the legs of their remounts as if the horses were coins. And the Sidhi – the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.

When you walk the Way of Silk, walk with empty hands. The last cup of water is never yours to keep. The ninth cord is never cut. And the amulet Kalrantha does not grant wishes – it only watches you decide.

--- \textit{Dusana of the Raven Road, at a caravanserai whose name she refused to give. The ink was cut with rain.}
\end{lorebox}


\end{document}