\documentclass[11pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{a5paper, margin=1in}
\usepackage{setspace}
\onehalfspacing
\usepackage{microtype}
\usepackage{ebgaramond}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{array}
\usepackage{tabu}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=black, citecolor=black, urlcolor=gray}

\titleformat{\chapter}[hang] 
{\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{0pt}{-20pt}{20pt}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[RE]{\textit{\leftmark}}
\fancyhead[LO]{\textit{\rightmark}}
\renewcommand{\headrulewidth}{0.4pt}

\newenvironment{lorebox}{\begin{quote}\small\itshape}{\end{quote}}

\begin{document}

\frontmatter
\begin{titlepage}
\centering
\vspace*{2cm}
{\huge\bfseries Dhahara\par}
\vspace{0.5cm}
{\Large Throne of Brass\par}
\vspace{2cm}
{\large A Fate's Edge Expansion\par}
\vspace{1.5cm}
\vfill
\end{titlepage}

\tableofcontents

\mainmatter

\chapter*{Preface: The Two Ledgers of Dhahara}
\addcontentsline{toc}{chapter}{Preface: The Two Ledgers of Dhahara}

Dhahara is not a nation. It is an argument.

In the north, the mountain passes are held by Aeler viceroys who speak a creole of their own tongue, who ride elephants in silver howdahs, who keep fire shrines in the old Dhaharan way. Their ledgers are kept in slate and sealed with keystone wax. They call themselves the \textbf{Mountain Satrapies}, and they answer to the High King beneath the peaks -- but High Kings are distant, and tax collectors are not.

In the south, the Brass Coast is a patchwork of ports ruled by the \textbf{Kahfagian Admiralty}. Here, spice warehouses rise like cathedrals, and the navy patrols the monsoon lanes. The Admiralty styles itself a civilising hand; the Dhaharan call it a velvet glove over an iron hook. The Sidhi -- coastal cousins of the Ashaani -- are most numerous here, and they remember every slave raid.

Between them lies a no‑man's land of contested plains, ruined watchtowers, and the great river that some call the boundary and others call a wound. Dhahara has been split for so long that few remember a time when it was whole. The Mountain Viceroy and the Admiral‑Governor meet once a year on a bridge that is neither north nor south, exchange gifts that are also insults, and then return to their ledgers.

Yet Dhahara is far more than the sum of its foreign masters. Beyond the passes and the coast lie independent hill kingdoms, ancient temple‑cities, and riverine principalities where native dynasties still rule. Dhaharan merchants control the inland spice routes; Dhaharan priests maintain the great fire shrines; Dhaharan poets compose epics that the Aeler viceroys themselves commission for their courts. The ledger of Dhahara has two columns, but the ink is mixed with Dhaharan blood, and the interest is paid in Dhaharan grain.

This expansion presents Dhahara as a land of contradictions: monsoon wealth and wartime poverty, elephant cavalry and steam‑driven ropeways, fire‑worshippers and ancestor‑spirit cults. It is also the home of the \textbf{Kalrantha} amulet -- the green glass shard that an Ashaani prince stole, that cracked the world, and that turned a fertile kingdom into the Red Desert. Now the amulet has resurfaced, and the race to return it will tear open every old wound.

When you play in Dhahara, you will bargain with a Viceroy's chamberlain who weighs your words on golden scales. You will dodge the spies of the Admiralty's "Trade Protection Bureau". You will walk the monsoon‑shrouded decks of a Sidhi smuggler's dhow. And if you are lucky, you will reach the Himadri temple before the amulet's whispers turn you into the very thing you sought to contain.

\clearpage
\chapter{The Land and the Monsoon}

\section{Geography of the Split}

Dhahara is defined by its two halves and the seasonal wind that connects them to the wider world. But those two halves are only the zones of foreign administration. Beyond them lie many other Dhaharas.

\begin{itemize}
    \item \textbf{The Mountain Satrapies (North):} Aeler‑held highlands, terraced for tea and rice. Fortified passes, fire shrines on every peak, and a bureaucracy that measures time in candle‑marks. The Viceroy resides at \textbf{Sarvistan}, a city of domed libraries and elephant stables.
    \item \textbf{The Brass Coast (South):} Kahfagian‑administered ports along the Gulf of Oshiira and the eastern ocean. \textbf{Port Shed} is the largest, a maze of ropeways, spice auctions, and the Admiralty's notorious "Maritime Court". The further east you go, the more the coast blends into Ayokhan trading towns and Sidhi fishing villages.
    \item \textbf{The Scarred Plain:} A buffer of dry grassland, dotted with ruined watchtowers from the last war. Caravans cross it in one day if the weather holds, but bandits and tax inspectors cross it in hours.
    \item \textbf{The Himadri Peaks:} The northernmost spine, where the Aeler passes lead to the high plateau and the secret temple that once held the Kalrantha amulet. The temple is now a ruin, guarded by monks who speak a dialect no one else remembers.
    \item \textbf{The Dhaharan Heartlands:} Beyond the Satrapies and the Admiralty's reach lie a dozen independent principalities -- the hill kingdoms of the eastern ghats, the riverine federation of the middle Silver Divide, the desert trading cities of the west, and the temple‑states of the southern plateau. Each has its own royal house, its own patron deity, its own courts and coinage. The foreign powers tax what moves between them; they do not rule them.
\end{itemize}

\section{The Monsoon Trade}

The summer monsoon blows from the southwest, bringing ships from Ayokha, Taharka, and the distant eastern isles. The winter monsoon blows from the northeast, carrying Dhaharan silver and Aeler steel back to Kahfagia and Oshiira. This rhythm is the heartbeat of Dhaharan wealth -- and the clock that every faction watches.

Use the \textbf{monsoon clock [8]} to track the season. At 2, the first ships from Ayokha arrive; markets are flush, prices drop. At 4, the Admiralty raises tariffs. At 6, the northeast winds begin; merchants race to load. At 8, the window closes; any ship still in port is stuck for half a year.

\clearpage
\chapter{Peoples of Dhahara}

\section{The Dhaharan Heartlands (Majority Native Cultures)}

\subsection*{Pillars}
Elephant cavalry, temple cities, spice and indigo, ancestor shrines, monsoon festivals.

\subsection*{Voice}
Dhaharan speech is rich in proverbs and layered with honorifics. A farmer greets a merchant with "May the rain find your fields"; a merchant answers a prince with "May your treasury overflow". Insults are delivered through metaphor, and a well‑turned verse can end a feud.

\subsection*{Regional Variations}
The Dhaharan people are not a single nation but a mosaic of related cultures. The hill kingdoms of the east celebrate the \textbf{Feast of the First Rain} with elephant processions and flower‑strewn chariots. The river principalities of the central plains hold \textbf{stepwell courts} where disputes are settled at the water's edge. The desert cities of the west are famed for their \textbf{caravanserai poetry}, where traders compete in verse as they haggle over silk and salt. The southern plateau reveres \textbf{ancestral spirits} who speak through oracles at mountain shrines.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Offer a betel leaf} before serious negotiation -- refusal is a grave insult.
    \item \textbf{Never step over a threshold with your left foot first} -- the left is for the dead.
    \item \textbf{Share the first cup of tea with the eldest present} -- even if they are a servant.
\end{itemize}

\subsection*{Strings}
Ancestor tablet (proof of lineage), temple car token (right to a festival seat), weaver's loom weight (trade credit in the cloth markets).

\section{The Mountain Satrapies (Aeler--Dhaharan)}

\subsection*{Pillars}
Fire shrines, elephant cavalry, tea‑ledgers, mountain hospitality.

\subsection*{Voice}
Satrap officials speak in formal couplets, each line a weight and a counterweight. A joke is a ritual; a threat is a citation of ancient law.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Present a candle of the right color} -- white for petition, red for thanks, black for mourning. (Position +1)
    \item \textbf{Never refuse tea} -- the first cup is courtesy, the second is negotiation, the third is dismissal. (Breaking this ends all social rolls for the scene.)
\end{itemize}

\subsection*{Strings}
Fire‑shrine token (right to pray at a public shrine), tea‑ledger page (proof of credit), elephant‑road pass (priority in mountain passes).

\section{The Brass Coast (Kahfagian--Sidhi)}

\subsection*{Pillars}
Spice auctions, ropeways, maritime court, freedman's pride.

\subsection*{Voice}
Coast merchants speak with Kahfagian cadence but use Dhaharan honorifics. They are direct, suspicious, and quick to invoke the "Maritime Court" as a threat.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Declare your cargo's value twice} -- once to the clerk, once to the shrine. (The shrine version may be used as a String later.)
    \item \textbf{Never board a ship without a lamp} -- even at noon. (Lamp shows you are not a slaver.)
\end{itemize}

\subsection*{Strings}
Spice‑lot stamp (right to trade), ropeway pass (priority loading), freedman's seal (proof of emancipation from Ashaan).

\section{The Sidhi (Minority from Galanina)}

\subsection*{Pillars}
Sea trade, lighthouse courts, mixed ancestry (Vilikari and Ashaani), fierce independence. The Sidhi are a minority people who trace their origins to the ancient port city of Galanina, far to the west. Generations ago, many fled Ashaani slavery and settled along the Brass Coast. They remain a distinct community -- proud, insular, and extraordinarily skilled in navigation and trade.

\subsection*{Voice}
Sidhi speak a creole of Ashaani, Vilikari, and Dhaharan. They are laconic and watchful, with a dark humour about the Admiralty.

\subsection*{Etiquette Hooks}
\begin{itemize}
    \item \textbf{Light a lamp before landing} -- Position +1 in Sidhi homes.
    \item \textbf{Never ask about ancestry} -- the raids are a fresh wound. (Flips Audience from Warm to Cold.)
\end{itemize}

\subsection*{Strings}
Lighthouse charter, port‑weight (trade advantage), freedman's seal (honoured in Sidhi courts).

\section{The Imperial Factions (Outsiders with Interests)}

\subsection*{Oshiiran Grain Auditors}
Oshiira sends teams to monitor the grain trade; they are despised but necessary. Their Strings: grain ledger, weigh‑house receipt.

\subsection*{Ashaani "Trade" Envoys}
Veiled Sisters" agents, officially merchants, actually spying on the Brass Gate. Their Strings: spirit‑lamp, brass seal (counterfeit), a name that is not theirs.

\subsection*{Kahfagian Admiralty Officials}
The colonial administrators, naval officers, and customs clerks who run the Brass Coast. Their Strings: letter of marque, port writ, cipher book.

\subsection*{Ayokhan Merchant Houses}
Wealthy monsoon traders who rent warehouses and hire Sidhi crews. Their Strings: monsoon schedule, temple tithe stamp.

\clearpage
\chapter{The Free Principalities of the Heartlands}

\section{The Council of Nine Palanquins}
While the colonizers fight over the edges, the heart of Dhahara beats to its own rhythm. The \textbf{Council of Nine Palanquins} is a loose confederation of the most powerful native rulers. They meet only during the Eclipse Festival, carried on litter-borne thrones so their feet never touch the same soil as their rivals.

\begin{itemize}
    \item \textbf{The River Rajas:} Control the Silver Divide. They tax every boat that crosses the middle waters. Their capital is \textbf{Jalpur}, a city built on stilts above the flood line.
    \item \textbf{The Hill Maharajas:} Rule the eastern ghats. They control the tea terraces and the elephant breeding grounds. Their capital is \textbf{Girivan}, carved into a mountainside.
    \item \textbf{The Desert Sultans:} Control the western trade routes. They hold the monopoly on salt and glass. Their capital is \textbf{Registan}, a city of mirrors and wind towers.
    \item \textbf{The Temple States:} Religious enclaves that claim neutrality. They hold the ancient libraries and the fire shrines. Their capital is \textbf{Agnigarh}, the City of Eternal Flame.
\end{itemize}

\section{The Stepwell Network}
Information flows through Dhahara like water. The \textbf{Stepwell Network} is a system of spies, merchants, and priests who use the ancient stepped wells as dead drops. A message carved into a wet stone wall can travel from the desert to the coast in a week.

\textbf{Mechanic:} Players with the \textbf{Stepwell Contact} String can ask one question about local rumors per session. The answer is always true, but may be cryptic.

\section{The Spice Road Confederacy}
Independent merchants who refuse to pay the Admiralty's tariffs or the Satrapy's tolls. They move goods through the Heartlands using secret paths known only to their guild masters. They are heavily armed and fiercely protective of their neutrality.

\textbf{Mechanic:} Joining the Confederacy grants access to \textbf{Safe Houses} in every major Heartland city. However, breaking Confederacy law results in immediate exile and a bounty on the character's head.

\section{Cultural Mechanics: The Weight of Honor}
In the Heartlands, reputation is currency. The \textbf{Honor Clock [6]} tracks a character's standing in native society.
\begin{itemize}
    \item \textbf{Gain Honor:} Fulfilling oaths, protecting the weak, donating to temples.
    \item \textbf{Lose Honor:} Breaking promises, insulting elders, harming sacred animals.
    \item \textbf{Effect:} High Honor grants +1 Position in social encounters with natives. Low Honor makes natives hostile and prices double.
\end{itemize}

\clearpage
\chapter{Patrons of Dhahara}

\section{Regional Faces of Known Patrons}

\subsection{Mykkiel -- The Judge of the Peak}
In the Mountain Satrapies, Mykkiel is not a distant arbiter but a local magistrate. His temples are audience halls, his rites are public hearings. He demands that every oath be witnessed by two elders. The Veiled Sisters fear him because he has already tried them in absentia.

\textbf{Regional Rite -- The Unsealed Testimony:} Once per session, compel a spirit to speak truth for one question; the spirit's former master learns your face. (Mark +2 Obligation)

\subsection{Aveh -- The Dust‑Rider of the Scarred Plain}
Here, Aveh is a djinn -- a spirit of the sandstorm, the liberator of slaves, the patron of those who flee without paying. She appears as a rider without a face, offering an exit from any prison or contract. Her price is always the same: you must never close a door behind you.

\textbf{Etiquette:} When you flee, leave a coin at the threshold. (The GM may convert the first SB of pursuit into a "missed footing" complication instead of harm.)

\textbf{Rite -- The Unlatched Chain:} Break one physical bond (shackle, rope, locked door) as a free action. (Mark +1 Obligation; the broken object may later reform.)

\subsection{Ikasha -- The Lotus in the Well (Sidhi variant)}
Among the Sidhi, Ikasha is the dream that rises from dark water. She offers refuge to escaped slaves and spirits. Her shrines are wells with a single lamp floating on the surface.

\textbf{Rite -- The Black Water:} Once per journey, you may hide one object or person in a Sidhi well for a season; no one will look there. (Mark +1 Corruption; the well may ask for a story in return.)

\section{New Patrons of Dhahara}
\emph{The following patrons are worshipped across the Dhaharan heartlands, from the hill kingdoms to the desert cities. They have no single temple; their shrines are everywhere -- at fords, in marketplaces, under ancient banyan trees.}

\subsection{Vivas -- The Unblinking Ledger}
\paragraph{Domain:} The sun, audits, the moment of truth.
\paragraph{Lore:} Vivas is a spirit of the high passes, where the sun bakes the stone and shadows are short. He does not speak; he witnesses. His followers are tax collectors, truth‑seekers, and those who fear the dark.
\paragraph{Favored Deeds:} Exposing a liar, paying a debt at noon, climbing to a shrine.
\paragraph{Taboos:} Lying in direct sunlight, hiding a tax receipt, eating before noon.
\paragraph{Low Rite -- The Shadowless Hour (4 XP):} For one scene, you cannot be deceived by visual illusions. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Vivas's Gaze (8 XP):} Target must tell one truth about themselves or suffer --1 die to all rolls until dawn. (Mark +2 Obligation)
\paragraph{High Rite -- The Unforgiving Ledger (14 XP):} Name a debt; the debtor cannot sleep until they pay. (Mark +3 Corruption; the debt may be spiritual, not monetary.)

\subsection{Indu -- The Well That Reflects}
\paragraph{Domain:} Moon, hidden water, second chances.
\paragraph{Lore:} Indu is the spirit of the well where a Sidhi child hid from slavers. She is kind, cautious, and forgetful -- but she remembers cruelty. Her followers are smugglers, midwives, and the keepers of way‑stations.
\paragraph{Favored Deeds:} Hiding a fugitive, drawing water at night, whispering a secret to the moon.
\paragraph{Taboos:} Using a well to drown evidence, silencing a child's cry, reflecting a lie in water.
\paragraph{Low Rite -- The Moon Pool (4 XP):} Look into any reflective surface; see the nearest safe hiding place. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Indu's Cup (8 XP):} A waterskin becomes "memory" -- anyone who drinks from it recalls a forgotten fact. (Mark +2 Obligation)
\paragraph{High Rite -- The Second Well (14 XP):} Create a temporary well that lasts one night. The water heals 1 Harm but imposes a geas. (Mark +3 Obligation)

\subsection{Jvala -- The Three Fires}
\paragraph{Domain:} Hearth, forge, funeral pyre.
\paragraph{Lore:} Jvala is three spirits in one -- the cook fire, the smith's flame, and the cremation torch. He is the patron of cooks, blacksmiths, and undertakers. He has no opinion on war; he cares only about heat.
\paragraph{Favored Deeds:} Keeping a kitchen fed, forging a tool for a worthy cause, performing a proper funeral.
\paragraph{Taboos:} Letting a hearth go cold in winter, using a forge for weapons of assassination, burning a body without witness.
\paragraph{Low Rite -- The Smell of Smoke (4 XP):} For one scene, you know the location of the nearest fire. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Jvala's Seal (8 XP):} Touch a metal object; it becomes fire‑resistant for one day. (Mark +2 Obligation)
\paragraph{High Rite -- The Three Fires (14 XP):} Once per campaign, when you die, you may be cremated immediately. Your ashes grant a boon to your heirs (GM determines). (Mark +3 Corruption; you cannot be resurrected.)

\subsection{Marut -- The Wind That Writes}
\paragraph{Domain:} Monsoon, messages, the speed of news.
\paragraph{Lore:} Marut is a spirit of the monsoon wind, who carries rumours faster than horses. He is fickle, generous, and easily bored. His followers are messengers, spies, and sailors.
\paragraph{Favored Deeds:} Delivering a message through a storm, reading a wind‑chart correctly, warning a village of a flood.
\paragraph{Taboos:} Burning a letter without reading it, silencing a messenger, hoarding wind‑charms.
\paragraph{Low Rite -- The Written Breath (4 XP):} Whisper a message into a leaf; the wind will carry it to the nearest temple of Marut. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Marut's Chart (8 XP):} For one leg, your ship or caravan gains +1 Speed. (Mark +2 Obligation)
\paragraph{High Rite -- The Monsoon's Secret (14 XP):} Once per season, you may predict a storm's path with perfect accuracy. (Mark +3 Corruption; the prediction must be shared freely.)

\subsection{Bhumi -- The Elephant That Remembers}
\paragraph{Domain:} Ancestors, memory, the weight of tradition.
\paragraph{Lore:} Bhumi is a spirit of the elephant herds that once roamed the Dhaharan plains. He is slow, patient, and never forgets a face. His followers are cavalrymen, archivists, and those who keep family ledgers.
\paragraph{Favored Deeds:} Recording a family history, protecting a herd, eating an elephant's favourite fruit.
\paragraph{Taboos:} Selling elephant ivory, forgetting a debt to a relative, riding a pregnant elephant.
\paragraph{Low Rite -- The Stomp of Memory (4 XP):} For one scene, you gain +1 die to recall any fact about Dhaharan history. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Bhumi's Seal (8 XP):} Touch an object; you know the name of its maker. (Mark +2 Obligation)
\paragraph{High Rite -- The Herd's Witness (14 XP):} Summon an ancestor spirit for one question; the spirit may lie if it is protecting a secret. (Mark +3 Corruption)

\subsection{Nadi -- The River That Divides}
\paragraph{Domain:} Boundaries, fords, the space between north and south.
\paragraph{Lore:} Nadi is the spirit of the great river that splits Dhahara. She is neutral, patient, and merciless to those who try to cross her without a proper ferry. Her followers are ferrymen, bridge‑wardens, and diplomats.
\paragraph{Favored Deeds:} Building a safe crossing, rescuing a drowning person, cleaning a riverbank.
\paragraph{Taboos:} Polluting the river, charging a corpse for passage, cutting a bridge without warning.
\paragraph{Low Rite -- The Ford's Price (4 XP):} Know the safest crossing point on any river. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Nadi's Witness (8 XP):} Any oath sworn on the river is witnessed by Nadi; breaking it imposes --2 dice to all social rolls until atonement. (Mark +2 Obligation)
\paragraph{High Rite -- The Divided Waters (14 XP):} Once per season, you may part a narrow river for one minute. (Mark +3 Corruption; the river may ask for a sacrifice in return.)

\subsection{Vetra -- The Silence of the Blade}
\paragraph{Domain:} Justice, execution, the end of argument.
\paragraph{Lore:} Vetra is a spirit of the old Dhaharan courts -- the executioner's sword that fell on oath‑breakers. She is cold, final, and without mercy. Her followers are judges, assassins, and those who have been wronged beyond forgiveness.
\paragraph{Favored Deeds:} Carrying out a lawful execution, protecting a witness, destroying a false idol.
\paragraph{Taboos:} Drawing the blade without a witness, killing a prisoner who has surrendered, forging a death warrant.
\paragraph{Low Rite -- The Unsheathed (4 XP):} For one scene, your melee attacks gain +1 die against oath‑breakers. (Mark +1 Fatigue)
\paragraph{Standard Rite -- Vetra's Judgement (8 XP):} Name a target; for one day, any lie they tell will be exposed by a chill wind. (Mark +2 Corruption)
\paragraph{High Rite -- The Final Argument (14 XP):} Once per campaign, you may kill a mortal oath‑breaker with a single blow -- but you must also destroy their name from all records. (Mark +3 Corruption; you become a fugitive from the victim's family.)

\subsection{Shri -- The Random Coin}
\paragraph{Domain:} Luck, windfall, the unexpected profit.
\paragraph{Lore:} Shri is a spirit of the marketplace -- the coin that is found, the trade that is offered by a stranger, the spice shipment that arrives early. She is capricious, kind to beggars, and cruel to misers.
\paragraph{Favored Deeds:} Giving alms, playing dice for a friend's debt, sharing a surprise profit.
\paragraph{Taboos:} Hoarding a found coin, rigging a game of chance, refusing to gamble when the stakes are low.
\paragraph{Low Rite -- The Found Penny (4 XP):} You find a single coin on the ground. It is always exactly the amount you need for a small purchase. (Mark +1 Fatigue; the coin vanishes after use.)
\paragraph{Standard Rite -- Shri's Smile (8 XP):} Once per session, you may reroll a failed trade or negotiation roll. (Mark +2 Obligation)
\paragraph{High Rite -- The Floating Market (14 XP):} For one hour, you may buy or sell any mundane item in any settlement, regardless of inventory. (Mark +3 Corruption; prices are set by the GM, not market logic.)

\clearpage
\chapter{The Kalrantha Amulet}
\label{chap:kalrantha-adventure}

\section{What It Is}
A small, tear‑shaped amulet of green glass, set in a brass frame that never tarnishes. Warm to the touch when someone nearby is about to decide. It hums with the element of \textbf{Fate} — not the fate of prophecies, but the cold gravity of consequence. Like a stone falling, the amulet does not care whether the landing is soft or hard. It only knows that the fall must end.

\section{True History (GM Only)}

The amulet was created by the priests of the Himadri temple, high in the peaks that separate Dhahara from the eastern steppes. They were not trying to make a weapon. They were trying to contain a demon of indecision — a spirit that fed on the hesitation between choices, that grew fat on the breaths held before a verdict. They bound it into a shard of green glass and sealed it beneath a stone where no sun could reach.

For centuries, it slept.

Then an Ashaani prince, whose name has been struck from every record, learned of the amulet. He was a younger son with no inheritance, a man whose ambition had curdled into a hunger for something — anything — that would make the world remember him. He carved a bloody path across the Way of Silk, through Dhaharan hill country and Aeler passes, until he reached the Himadri temple. He slew the monks. He cracked the seal. And he took the amulet for himself.

He brought it back to his kingdom in Ashaan.

The amulet did not grant wishes. It whispered. It offered perfect confidence — the absolute certainty that every choice was right, every path ordained, every consequence deserved. The prince believed himself invincible. He raised armies. He conquered cities. He burned temples that had stood for a thousand years. And with every decision, the demon inside the glass drank a little more of his soul.

The prince's kingdom became the Red Desert. Not overnight, but over a single generation. The rivers dried. The grass turned to dust. The people who did not flee died of thirst or madness. Some say the prince built a lavish tomb for himself as he was dying. Others claim he was last seen walking into a sandstorm, still wearing the amulet, still smiling.

The amulet was lost. The desert remained.

Now it has resurfaced in Dhahara, carried by a relic hunter who did not know what he had found. The demon is not dead — only sleeping. The temple’s wards are cracked. And the amulet wants to be used again.

\section{Game Mechanics (Full Rules)}

The amulet is a **Tier V artifact** with the following properties, known incrementally to players as the story unfolds.

\subsection*{Known to Players (Initial)}
Once per session, the wearer may reroll a single die after seeing the result. That is all.

\subsection*{Hidden Mechanics (GM Only)}

\begin{itemize}
    \item \textbf{Fate’s Gravity:} The reroll ability is the amulet’s smallest gift. Each time it is used, the GM marks a segment on a hidden \textbf{Fate Corruption Clock [6]}. When the clock fills, a major twist occurs (see \textit{Corruption Manifestations} below).
    \item \textbf{Anti‑Magic Aura (Near):} Any spell or magical effect that directly alters probability, fate, or causality (e.g., a Rite that forces a reroll, a blessing that guarantees success, or a curse that imposes automatic failure) suffers \textbf{+2 DV} within Near range of the amulet. The wearer is immune.
    \item \textbf{Story Beats Become Gravity:} When the wearer rolls one or more \textbf{1s}, those Story Beats are not added to the GM’s pool. Instead, they are transferred to the \textbf{next player character who acts}. The GM must spend them on complications that affect that character’s action, and those complications must be \textit{inevitable} in tone (e.g., “the floor gives way,” “your sword strikes a hidden chain,” “you recognise your own face in a reflection – and it is weeping”). Unspent beats vanish after that action.
    \item \textbf{Obligation to the Demon:} Each time the amulet’s reroll is used, mark \textbf{+1 Obligation} (treat as a Terrestrial Patron with Capacity = wearer’s Spirit). When Obligation exceeds Capacity, the wearer has nightmares of falling; the GM gains 1 SB per excess segment. When Obligation reaches double Capacity, the demon fragment manifests as a voice whispering a single suggestion (“Take the left path,” “Trust the merchant,” “Draw your sword”). The wearer may ignore it, but doing so clears all Obligation and the amulet becomes inert for one session.
\end{itemize}

\subsection*{Corruption Manifestations (d6, GM rolls when Fate Corruption Clock fills)}
\begin{enumerate}
    \item A sudden sandstorm traps the party. Lose 1 Supply and mark Fatigue 1.
    \item A character’s trusted ally betrays them (the demon planted the suggestion).
    \item A forgotten debt surfaces – someone the party wronged appears with a writ.
    \item The amulet glows; a nearby spirit is drawn to it. The spirit is not hostile, but it asks a question that has no good answer.
    \item The ground trembles; an old collapse seals a path. The party must find another way.
    \item The demon speaks aloud for one sentence. The sentence is a secret about a PC. The PC takes 1 Stress (Harm 1, mental).
\end{enumerate}

\section{Showpiece Adventure: Return of the Kalrantha}

\subsection{The Setup}
An archaeologist — a Dhaharan scholar named \textbf{Vikram of the Broken Wheel} — has found the amulet in the Ashaani desert. He knows its history. He wants to return it to the Himadri temple before the demon wakes. He is dead within the first scene, poisoned by a Veiled Sister’s agent. The party is hired by his Tulkani assistant to finish the job.

\subsection{The Race}
Four factions want the amulet. Each will approach the party with an offer, a threat, or a favour. The party may ally with none, one, or try to play them against each other.

\begin{longtable}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.3\textwidth}|}
\hline
\textbf{Faction} & \textbf{Agent} & \textbf{Offer} \\
\hline
Ashaani (Veiled Sisters) & Nadia, a charming merchant & A fortune in gold (10 Measures) and a writ of safe passage in Ashaan. \\
\hline
Oshiiran Grain Auditors & Mei‑Lin, cold and precise & Safe passage through Oshiiran waters and a letter of credit (5 Measures). \\
\hline
Kahfagian Admiralty & Captain Red‑Hook, cynical privateer & A ship (a fast cutter) and a letter of marque. \\
\hline
The True Believer & Brother Ananda, silent monk & Only the truth: the amulet will kill its user. He has no coin, but he knows the temple’s wards. \\
\hline
\end{longtable}

\subsection{Key Scenes}

\paragraph{1. The Poisoned Scholar (Ashaan bazaar)}
Vikram dies in a crowded market, his last words: “The ninth knot…” His Tulkani assistant (a woman named \textbf{Sarnai}) gives the party the amulet and a map to the Himadri temple. The murderer — a Veiled Sister’s agent in a servant’s dress — watches from a rooftop. If confronted, she flees.

\paragraph{2. The Three Offers (Port Shed, Sidhi lighthouse)}
The party arrives at a Sidhi port. Each faction approaches separately. Brother Ananda finds them at dawn, offering only a warning. The others offer bribes. The party may accept multiple offers, but doing so creates Debt Clocks.

\paragraph{3. The Scarred Plain (Dhaharan overland)}
The party crosses the Red Desert’s edge. If they trusted one faction, that faction betrays them (or is betrayed). A sandstorm (courtesy of Aveh) provides cover. The party must make a group Survival roll (DV 4, Desperate). On a failure, they lose 2 Supply and one PC gains a \textit{Lost} Condition (disoriented, –1 die to navigation).

\paragraph{4. The Mountain Pass (Aeler toll station)}
The Aeler viceroy demands a toll: “A story, a secret, or a sack of silver.” The party must choose. If they pay in story, the amulet trembles – the demon is listening. Tick the Fate Corruption Clock once.

\paragraph{5. The Himadri Temple (climax)}
The temple is a ruin. The central altar is a stone table with a depression shaped like the amulet. Brother Ananda performs a ritual that requires the amulet to be placed on the altar. The ritual takes five rounds. Each round, a wave of attackers arrives (Ashaani assassins, Oshiiran mercenaries, Kahfagian sailors – or all three, if the party angered everyone).

Use the following clock to track the ritual:
\begin{center}
\textbf{Ritual Completion [5] – advance by 1 at the end of each round.}
\end{center}
Each round, the GM may introduce a complication (choose or roll d4):
\begin{enumerate}
    \item A sandstorm sweeps through; all ranged attacks at Desperate.
    \item The amulet glows; one PC hears the demon’s whisper. They must roll Resolve (DV 4) or lose their action.
    \item A section of the temple collapses; split the party.
    \item The attacker waves include a Veiled Sister’s handmaiden (Cap 3 adversary).
\end{enumerate}

\paragraph{6. The Choice}
When the ritual reaches 5 segments, Brother Ananda cries: “Now! Place it on the altar!” The amulet is placed. The seal closes. But for one moment, the demon is free – and it offers the wielder a wish.

\begin{quote}
\textit{“Speak your desire. I am the gravity of choice. I will make it so.”}
\end{quote}

The player may take any wish – wealth, power, a dead ally returned, a foe destroyed. If they do, the demon is not sealed; it escapes into the world. Brother Ananda dies (his ritual consumed him). The amulet becomes inert glass.

If they refuse, the demon howls and is pulled back into the amulet. The seal holds. Brother Ananda collapses, alive but aged. He whispers: “You have bought the world time. But it will whisper to you now – every night.” The PC who refused gains the condition \textbf{Demon’s Whisper} (see below).

\subsection{Twists and Complications}

\begin{itemize}
    \item \textbf{Sarnai’s Secret:} Vikram’s assistant is actually a Black Hand agent (Ashaani) in disguise. She has been guiding the party to remove the other factions. She reveals herself at the temple, trying to take the amulet. The party may fight or persuade her.
    \item \textbf{Mei‑Lin’s True Mission:} The Oshiiran agent is not an auditor but a Narrow Captain who wants to start a war. She plans to use the amulet as a pretext. If she survives, she becomes a recurring antagonist.
    \item \textbf{Captain Red‑Hook’s Vengeance:} The Kahfagian captain is a Sidhi who lost his family to the Admiralty. He plans to sell the amulet to Ayokha. He is not evil – only desperate. The party may redeem him with a promise.
    \item \textbf{Brother Ananda’s Past:} He is not a monk. He is the last surviving priest of the original Himadri temple, kept alive by the amulet’s residual magic. If the amulet is sealed, he will age to his true years and die within a week.
\end{itemize}

\subsection{Resolution and Aftermath}

\paragraph{If the amulet is sealed:}
The party gains a permanent boon: \textbf{The Demon’s Debt}. Once per session, any party member may reroll a single die, but the GM gains 1 SB (the demon still stirs). Brother Ananda dies peacefully. The amulet is locked in the temple’s vault. The factions disband – for now.

\paragraph{If the amulet is destroyed:}
Brother Ananda dies in the destruction. The party gains a curse: \textbf{The Unsealed Vow}. The demon is not bound; it will hunt them in their dreams. At the start of each session, the GM rolls a Dread Clock [4]. On a 1, a nightmare manifests as a complication: a betrayal, a lost memory, or a physical wound (1 Fatigue). This curse can only be lifted by a quest to bind the demon permanently.

\paragraph{If the wish is used:}
The demon is freed. The amulet becomes a green pebble. Brother Ananda is slain. The party must now hunt the demon, which can possess mortals and twist their choices. This becomes a new campaign arc. The demon’s first act is to possess the character who made the wish, unless they succeed on a Spirit + Resolve test (DV 6). On success, they resist – but the demon marks them: they now have the \textbf{Demon’s Debt} as above, and the demon will return.

\subsection{Running the Demon (If Freed)}
\begin{itemize}
    \item The demon’s name is \textbf{Varash the Unmaker}. It cannot be killed – only bound.
    \item It manifests as a voice, a shadow, or a cold draft. It can possess mortals for short periods.
    \item Its goal is to create chaos: to push rulers into war, lovers into betrayal, merchants into ruin.
    \item To bind it again, the party must retrieve the original sealing rites from the ruins of the Himadri temple and perform them at the site of the prince’s death (the Red Desert). This is a full campaign.
\end{itemize}

\section{Adventure Hooks (Drop‑In)}
\begin{itemize}
    \item \textbf{The Missing Priest:} Brother Ananda has vanished. His last message is a single word: “Kalrantha.” The party must find him before the Veiled Sisters do.
    \item \textbf{The Demon’s Dream:} One PC wakes with a perfect memory of the amulet’s creation. They know where a fragment of the original seal is buried – but the dream was sent by Varash.
    \item \textbf{The Prince’s Tomb:} A sandstorm exposes a tomb in the Red Desert. Inside is a sarcophagus carved with nine knots. The amulet is not there – but a map to it is.
\end{itemize}

\begin{lorebox}
“Fate is not a path. It is the weight of the ground when you stop falling. The amulet does not guide; it only ensures. And the demon does not lie – it simply omits the interest.”

--- \textit{Dusana of the Raven Road, on the Kalrantha Amulet}
\end{lorebox}

\clearpage
\chapter{Adventure Framework}

\section{Adventure Seeds (Ten Drop‑Ins)}
\begin{enumerate}
    \item \textbf{The Elephant's Ledger:} A Mountain Satrapy official is selling stolen elephant‑road passes. The party must audit his accounts without revealing their true purpose.
    \item \textbf{The Smell of Monsoon:} A Sidhi crew is accused of smuggling spice; the real cargo is escaped slaves. The party must get them to a safe house.
    \item \textbf{The Forged Seal of the Admiralty:} A forger is creating false letters of marque. The Admiralty hires the party to find him; the forger is a former slave.
    \item \textbf{The Fire That Wouldn't Die:} A fire shrine in the mountains has been burning for a year without fuel. The priest is drawing on a deep vein of magic -- and something is answering.
    \item \textbf{The River Court:} Nadi's spirit is offended; the river has risen and flooded a village. The party must arbitrate between the north and south banks.
    \item \textbf{The Random Coin:} Shri appears as a beggar and offers the party a single coin. If they spend it wisely, they gain a fortune; if they hoard it, they lose everything.
    \item \textbf{The Silence of the Blade:} A judge is executing the wrong man. The party must prove the real killer's guilt -- or perform the execution themselves.
    \item \textbf{The Three Fires:} A smith has forged a sword that burns with Jvala's flame. He wants to sell it to the Viceroy; the Admiralty wants to steal it; the party is caught in the middle.
    \item \textbf{The Wind That Writes:} A monsoon message is intercepted. It contains a treaty that would end the cold war. The party must deliver it before the signatories are killed.
    \item \textbf{The Well That Reflects:} A Sidhi girl is hiding in a well; the Black Hand is searching for her. The party must get her to the coast safely.
\end{enumerate}

\section{GM Toolkit}

\subsection{Weather and Hazards (d12)}
\begin{enumerate}
    \item Monsoon squall -- all sails Desperate for a leg.
    \item Dust storm -- Vision --1 Position.
    \item Landslide -- mountain pass closed; detour adds 2 segments.
    \item Elephant herd crossing -- delay 1 day; harm if startled.
    \item River flood -- ferry destroyed; lose 1 Supply.
    \item Earthquake -- temple ruins shift; new paths open.
    \item Smoke from a forest fire -- test Endurance or mark Fatigue.
    \item Perfect calm -- no wind; ships drift.
    \item Meteor shower -- omens; the GM gains 1 SB.
    \item Ashfall -- from distant volcano; all actions Desperate.
    \item Leeches -- in the Scarred Plain's wet season; lose 1 Supply to treat.
    \item The Ninth Hour -- sunset twice as long; the GM may ask a player to describe an omen.
\end{enumerate}

\subsection{Factions of Dhahara}
\begin{tabu} to \linewidth {|X|c|c|c|X|}
\hline
\textbf{Faction} & \textbf{Influence} & \textbf{Stability} & \textbf{Crisis} & \textbf{Goal} \\
\hline
Mountain Satrapy & 6 & 5 & 4 & Keep the passes closed to the Admiralty \\
Brass Coast Admiralty & 6 & 4 & 6 & Monopolise spice trade \\
Sidhi Port Guilds & 4 & 3 & 7 & Free the coast from Ashaan \\
Ashaani Agents & 4 & 3 & 5 & Recover Kalrantha \\
Oshiiran Auditors & 5 & 5 & 3 & Stabilise grain prices \\
Ayokhan Merchant Houses & 5 & 5 & 4 & Profit from both sides \\
Council of Nine Palanquins & 5 & 6 & 5 & Maintain Heartland independence \\
Spice Road Confederacy & 4 & 4 & 6 & Protect trade routes from tariffs \\
\hline
\end{tabu}

\clearpage
\chapter{Dhahara — ``The Nine Kingdoms and the Brass Coast''}
\label{sec:dhahara}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Dhahara is a land of extremes: the snow‑capped Himadri peaks to the north, the burning central plateau of the Sarpies, and the humid, spice‑laden Brass Coast to the south. A patchwork of nine kingdoms — some ancient, some newly risen — squabble over river rights, elephant herds, and the favour of distant empires. The Aeler viceroys tax the mountain passes; the Kahfagian Admiralty polices the ports. But the heart of Dhahara is its own: a land of towering temples, forgotten iron mines, and kings who claim descent from the sun.

\vspace{2mm}
\textbf{What you notice first:} The smell of cardamom and dust. The way every village has a painted eye on its gate. The constant sound of bells — on cattle, on temple towers, on the ankles of dancers.

\textbf{The one rule that kills newcomers:} Never refuse a gift of betel leaf. It is an oath of hospitality. Refusing it is an insult that can end a bloodline.
\end{tcolorbox}

\subsection*{The Nine Kingdoms (Quick Reference)}
The nine kingdoms are a fluid concept: alliances shift, borders bleed, and a kingdom may dissolve into a collection of hill forts after a bad monsoon. The most influential are:
\begin{itemize}
    \item \textbf{Ahimsa (the Peacock Throne):} The richest, most stable kingdom. Its king is a patron of poets and a secret follower of the Covenant.
    \item \textbf{Vajrapur (the Iron Kingdom):} Controls the Himadri iron mines. Its elephants are armoured in steel.
    \item \textbf{Tamralipta (the Coast of Brass):} The main port, a Sidhi trading post and a Kahfagian client. Its harbour is a forest of masts.
    \item \textbf{Kalinga (the Eastern Delta):} A maze of rivers and mangrove isles. Its rulers are masterless pirates who pay tribute to no one.
    \item \textbf{Mahishmati (the Buffalo Throne):} A pastoral kingdom of the central Sarpies. Its warriors ride water buffalo and worship a horned god.
\end{itemize}

\subsection*{Prominent NPCs of Dhahara}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Maharaja Vikramaditya & King of Ahimsa & Unite the nine kingdoms under the Peacock Throne & He has a pet tiger that follows him everywhere; the tiger is drugged. \\
\hline
Rani Padmavati & Queen of Vajrapur & Protect the iron mines from Aeler encroachment & She wears a necklace of Aeler toll‑keys – each is a vanquished viceroy. \\
\hline
Syndic Olena & Kahfagian port‑master of Tamralipta & Maximise tariff revenue & She keeps a ledger in her own blood; she says ink fades. \\
\hline
Brother Haridas & Covenant judge of the inland circuit & Impose law on the lawless & He has never drawn a blade; his voice alone has stopped wars. \\
\hline
Captain Suleiman & Sidhi lighthouse commander & Keep the Brass Coast free of Ashaani slavers & He lost his nose to a shark; he wears a brass replacement. \\
\hline
\end{longtable}

\section{The Brass Coast — ``The Shore of Silver and Spice''}
\label{sec:brass-coast}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere (Brass Coast)}]
The Brass Coast is Dhahara’s southern rim: a humid, feverish coastline of palm groves, salt pans, and crowded port cities. The Kahfagian Admiralty administers it as a “client province,” but the Sidhi run the real trade, and the local rajas chafe under foreign tariffs. The air smells of pepper, cinnamon, and the low tide of a thousand ships. Here, a man can buy anything – a ship, a slave, a prophecy – and sell anything except his own shadow.

\vspace{2mm}
\textbf{What you notice first:} The heat pressing down like a damp cloth. The constant noise of haggling in a dozen languages. The way the sea is the same colour as the sky, so the horizon disappears.

\textbf{The one rule that kills newcomers:} Never board a ship that has a green lantern at the mast. It means the captain is under a death vow – and you are the payment.
\end{tcolorbox}

\subsection*{Names of the Brass Coast}
Inspired by the southern Indian and Sri Lankan coasts: Tamil, Sinhalese, and Sidhi influences. Surnames often denote a port, a boat, or a trade.

\begin{longtable}{|p{0.2\textwidth}|p{0.2\textwidth}|p{0.25\textwidth}|p{0.25\textwidth}|}
\hline
\textbf{Male} & \textbf{Female} & \textbf{Surname} & \textbf{Epithet} \\
\hline
Kannan & Ilavarasi & Saliya & the Tide‑Reader \\
\hline
Rajendran & Mangai & Pattinam & Brass‑Fingered \\
\hline
Siddharth & Anjali & Chettiar & Coin‑Counter \\
\hline
Arjunan & Nila & Pallava & Lighthouse Born \\
\hline
Muthu & Kalyani & Mudaliar & Shark‑Tooth \\
\hline
\end{longtable}

\section*{Spades — Places (Brass Coast)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Tamralipta Harbour} — A forest of masts; the smell of tar and spices. \textit{The harbour is a maze of floating houses; the residents are all Sidhi. They never touch the shore.}
\item \textbf{Pepper Warehouse} — A long building with slatted walls; the air is sharp. \textit{The workers are slaves; their eyes are red from pepper dust. They never blink.}
\item \textbf{Lighthouse of the Brass Flame} — A stone tower with a copper mirror. \textit{The light is green; it shows the faces of the drowned. The keeper knows their names.}
\item \textbf{Salt Pan of the Silent Widows} — A flat expanse of white crust; women in black work the rakes. \textit{The women are all widows; their husbands were lost at sea. The salt is bitter.}
\item \textbf{Kahfagian Admiralty House} — A white building with a veranda; a flag with a lantern. \textit{The flag is lowered at sunset; the Admiral is drunk by dawn.}
\item \textbf{Sidhi Court of the Lighthouse} — A circular chamber with a brass scale. \textit{The judge is blind; she hears truth in the echo of a coin.}
\item \textbf{Palm Grove of the Elephants} — A grove where temple elephants are bathed. \textit{The elephants are painted with turmeric; the paint is a prayer.}
\item \textbf{Coral Reef of the Sunken Ship} — A reef just offshore; a mast breaks the surface. \textit{The ship was a slaver; the slaves escaped. They live in the reef now.}
\item \textbf{Pepper Auction Floor} — A crowded hall; voices bid in claps. \textit{The auctioneer has a hammer of whalebone. It cracks when he lies.}
\item[J] \textbf{The Green Lantern Tavern} — A low building with a green lantern. \textit{The lantern is never lit. The owner says it would attract the wrong kind of dead.}
\item[Q] \textbf{The Shipbreaker’s Yard} — A beach where old ships are dismantled. \textit{The workers are convicts; they wear chains that drag in the sand.}
\item[K] \textbf{The Admiral’s Palace} — A former raja’s fort, now occupied by Kahfagian officers. \textit{The flag is flown upside down; the Admiral does not know the difference.}
\item[A] \textbf{The Undercliff} — A smuggler’s cove accessible only at low tide. \textit{The entrance is a cave; the cave leads to a hidden harbour. The harbour is a city of thieves.}
\end{enumerate}

\section*{Hearts — People \& Factions (Brass Coast)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Sidhi Lighthouse Keeper} — Blind in one eye; carries a brass telescope. \textit{He can see the future through the lens – but only the futures that end at sea.}
\item \textbf{Kahfagian Port‑Master} — A fat man with a ledger and a sword. \textit{He has two ledgers; one is for bribes, one is for the Admiral. He has confused them twice.}
\item \textbf{Pepper Factor} — A thin woman with red‑stained fingers. \textit{She can tell the quality of pepper by taste; she has never been wrong. Her tongue is black.}
\item \textbf{Temple Elephant} — A painted elephant with a mahout on its neck. \textit{The elephant can speak – one word per year. It has not spoken in a decade.}
\item \textbf{Shipbreaker} — A muscular man with a hook for a hand. \textit{He lost his hand to a cursed ship. The ship is still in the yard. It creaks at night.}
\item \textbf{Smuggler Captain} — A woman with a parrot on her shoulder, a curved knife on her belt. \textit{The parrot is her navigator; it knows the reefs by sound.}
\item \textbf{Admiral of the Brass Coast} — A grey‑haired man with a glass eye. \textit{The glass eye is from a defeated pirate; the pirate still watches through it.}
\item \textbf{Raja of the Inland Hills} — A young man in silk, come to the coast to sell iron. \textit{He has never seen the sea. He is afraid of it.}
\item \textbf{Ashaani Refugee} — A family hiding in a pepper warehouse. \textit{They paid the factor to forget them. The factor forgets nothing.}
\item[J] \textbf{The Drowned Priest} — A ghost in a wet robe who walks the shore at low tide. \textit{He asks for a coin to pay the ferryman. Give it, and he will not follow.}
\item[Q] \textbf{The Black Bishop} — A Covenant judge who rules on maritime disputes. \textit{He wears a ring made from a ship’s nail; the ship was the first slave ship he freed.}
\item[K] \textbf{The Green Lantern’s Owner} — A woman with no shadow. \textit{She sold her shadow for the tavern. The shadow now serves drinks.}
\item[A] \textbf{The Sea King of Kalinga} — A pirate lord with a fleet of painted dhows. \textit{He has nine wives; each is from a different kingdom. The ninth is a spy.}
\end{enumerate}

\section*{Clubs — Complications/Threats (Brass Coast)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Pepper Dust Fire} — A spark ignites a warehouse; the air is flammable. \textit{The workers are trapped; the party must choose – save them or save the cargo.}
\item \textbf{Admiral’s Press Gang} — Kahfagian marines grab bystanders for the fleet. \textit{The marines are drunk; they do not recognise rank.}
\item \textbf{Green Lantern Lit} — Someone lights the tavern lantern. \textit{The dead come. They are looking for the person who killed them.}
\item \textbf{Shipbreaker’s Curse} — The cursed ship in the yard breaks its chains. \textit{It wants a new victim. The victim must be a ship’s captain.}
\item \textbf{Tariff Hike} — The port‑master doubles duties on all goods. \textit{The hike is a bribe; the bribe is to a Sidhi court. The court is not in session.}
\item \textbf{Coral Reef Piracy} — Pirates hide behind the reef; they attack at dusk. \textit{The pirates are escaped slaves; they do not take prisoners.}
\item \textbf{Elephant Rampage (Coast)} — A temple elephant breaks free and runs through the market. \textit{It is looking for its mahout, who was arrested. The mahout is innocent.}
\item \textbf{Smuggler’s Betrayal} — The smuggler captain sells the party to the Admiralty. \textit{The betrayal was a test; the party must prove they can be trusted – by killing her.}
\item \textbf{Sea King’s Tax} — A Kalingan pirate fleet blockades the harbour. \textit{The tax is one ship. Choose which one.}
\item[J] \textbf{The Drowned Priest’s Demand} — The ghost appears during the day. \textit{He asks for a name. The name must be someone who owes a death.}
\item[Q] \textbf{The Black Bishop’s Verdict} — The judge declares the party’s cargo forfeit. \textit{The verdict is false; the bishop is bribed. The party must prove it.}
\item[K] \textbf{The Sea King’s Ninth Wife} — The spy reveals herself. \textit{She offers a deal: the Sea King’s treasure for safe passage. The treasure is a curse.}
\item[A] \textbf{The Undercliff Collapse} — A rockslide seals the smuggler’s cove. \textit{Inside are the party and a monster. The monster is a salt‑crocodile. It is blind.}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage (Brass Coast)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Pepper Merchant’s Seal} — A wax seal; present it to sell pepper at double price. \textit{The seal is from a dead factor; his ghost haunts the warehouse.}
\item \textbf{Lighthouse Keeper’s Lens} — A brass tube; look through it to see the nearest reef. \textit{The lens is also a key; it opens the lighthouse vault.}
\item \textbf{Salt Widow’s Blessing} — A bag of salt; throw it into the sea to calm a storm. \textit{The salt is from the widows’ pans; they have blessed it with their tears.}
\item \textbf{Admiral’s Pardon} — A signed document; cancels one crime. \textit{The Admiral signs it with a glass eye; the signature is a lie.}
\item \textbf{Sidhi Court Ruling} — A written judgment; win any maritime dispute. \textit{The ruling is valid only if witnessed by a lighthouse keeper.}
\item \textbf{Elephant’s Tooth} — A tusk; present it to any temple for shelter. \textit{The tusk is from a rogue elephant; the temple killed it. They will not refuse.}
\item \textbf{Shipbreaker’s Hook} — The hook of a dead shipbreaker; use it to command the cursed ship. \textit{The ship will obey one command – then it will try to drown you.}
\item \textbf{Smuggler’s Map} — A chart of hidden coves. \textit{The map is tattooed on a sailor’s back; the sailor is dead. The skin is preserved.}
\item \textbf{Sea King’s Pardon} — A flag of truce; fly it to pass Kalingan waters unharmed. \textit{The flag is a shirt; the shirt belonged to a dead pirate.}
\item[J] \textbf{The Green Lantern’s Flame} — A small lantern that burns green. \textit{Light it to summon the dead; they will answer one question. Then they will ask one.}
\item[Q] \textbf{The Black Bishop’s Ring} — A ring made from a ship’s nail; wear it to be immune to admiralty arrest. \textit{The nail is from a slave ship; the slaves’ ghosts protect you.}
\item[K] \textbf{The Sea King’s Ninth Wife’s Veil} — A silk veil; wear it to be invisible to pirates. \textit{The veil is cursed; each use ages you one year.}
\item[A] \textbf{The Undercliff’s Treasure} — A chest of Sidhi gold; worth 10 Measures. \textit{The gold is from a slaver; it is warm. It wants to go back to the sea.}
\end{enumerate}

\subsection*{Quick use notes (Brass Coast)}
\begin{itemize}
\item Draw until all four suits appear. Highest rank sets main Clock.
\item Diamonds are maritime tokens (seals, lenses, maps) that shift Position.
\item If any A appears, echo \textbf{green fire and salt} – ghosts that demand, ships that remember, and tides that wait.
\end{itemize}

\subsection*{Brass Coast Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Tide:} Toss a crust into the harbour. Ignore first \textit{Green Lantern Lit} penalty.
\item \textbf{Knot the Anchor Chain:} Tie a sailor’s knot in a rope. Cancel \textit{Admiral’s Press Gang} or take –1 die on your next escape roll.
\item \textbf{Name to the Lighthouse:} Whisper a dead captain’s name into the green flame. Gain +1 Effect when smuggling this scene, but mark Obligation to the Sidhi.
\end{itemize}

\section{The Himadri Foothills — ``The Iron Passes''}
\label{sec:himadri}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere (Himadri Foothills)}]
The Himadri foothills are the rugged transition between the Sarpies plateau and the towering peaks of the Aelerian mountains. Here, the air is thin and cold, the rivers run white with snowmelt, and the passes are measured in breath‑counts rather than miles. Iron mines scar the hillsides, worked by Dhaharan labourers and overseen by Aeler engineers. Elephant caravans haul ore down to the Brass Coast; mule trains carry silver and tea up to the mountain holds. The Aeler have established a viceroyalty here – not conquest, but a long lease. The local rajas pay tribute in iron and receive protection from lowland enemies. It is an uneasy peace, written in bell‑codes and toll‑ledgers.

\vspace{2mm}
\textbf{What you notice first:} The ringing of iron hammers and the lowing of elephants. The way every bridge has a dwarf toll‑keeper with a brass scale. The silence after a bell rings – the silence is for the mountain to answer.

\textbf{The one rule that kills newcomers:} Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one.
\end{tcolorbox}

\subsection*{Prominent NPCs of the Himadri Foothills}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Viceroy Thrain & Aeler governor of the Iron Passes & Maximise ore output for the High King & He wears a bell around his neck; it rings when he lies. \\
\hline
Rani Jaya & Local raja of the eastern hills & Balance tribute to Aeler with loyalty to her clan & She lost an arm to a mine collapse; the prosthetic is a pickaxe. \\
\hline
Mine‑Captain Korr & Aeler engineer & Keep the tunnels safe and the quotas met & He counts his breaths in eights; he has never reached nine. \\
\hline
Tarkan & Elephant caravan master & Move iron to the coast before the snows close the passes & He speaks to his elephants in Dhaharan; they answer in Aeler. \\
\hline
Sister Gaura & Covenant judge of the hill circuit & Resolve disputes between Aeler law and local custom & She carries a water‑clock that runs on tears – her own. \\
\hline
\end{longtable}

\subsection*{Names of the Himadri Foothills}
A mix of Aeler gutturals and Dhaharan vowels. Surnames may be Aeler clan names or Dhaharan hill titles.

\begin{longtable}{|p{0.2\textwidth}|p{0.2\textwidth}|p{0.25\textwidth}|p{0.25\textwidth}|}
\hline
\textbf{Male} & \textbf{Female} & \textbf{Clan/Surname} & \textbf{Epithet} \\
\hline
Thrain & Jaya & Ironfoot & the Bell‑Breaker \\
\hline
Korr & Gaura & Stonehand & Breath‑Counter \\
\hline
Borin & Meera & Vikram & the Elephant’s Friend \\
\hline
Halk & Sita & Aezelbron & Bridge‑Warden \\
\hline
Rurik & Laxmi & Thrainsdottir & Toll‑Eater \\
\hline
\end{longtable}

\section*{Spades — Places (Himadri Foothills)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Iron Mine of the Crying Face} — A deep shaft with a carved face above the entrance. \textit{The face weeps rust; the tears are iron oxide. The miners say the mountain is mourning its dead.}
\item \textbf{Aeler Toll Bridge} — A stone arch over a gorge; a dwarf with a scale sits at the far end. \textit{The toll is not in coin – it is in weight. One pound of ore, one pound of story.}
\item \textbf{Elephant Caravan Stop} — A walled enclosure with water troughs and hitching posts. \textit{The posts are carved with the names of elephants who died on the pass; their ghosts still wait.}
\item \textbf{Forgotten Shrine of the Ninth Breath} — A stone hut with a bell that has no clapper. \textit{Ring it by striking the side, and the mountain will answer – once.}
\item \textbf{Smelter’s Yard} — A smoky valley where ore is turned into pig iron. \textit{The workers wear masks of wet cloth; the masks are dyed red by the dust.}
\item \textbf{Aeler Counting House} — A low stone building with brass scales and iron ledgers. \textit{The ledgers are written in Aeler script; the margins are full of Dhaharan curses.}
\item \textbf{Landslide Scar} — A raw gash on a hillside, where a pass was buried last monsoon. \textit{The bodies were never recovered. Their ghosts light fires at night.}
\item \textbf{Raja’s Hill Fort} — A timber palisade on a rocky knoll; a flag with a tiger. \textit{The flag is torn; it was torn by the claws of a tiger that the raja killed with his bare hands.}
\item \textbf{Tea Plantation} — Terraced slopes with low bushes; workers pick leaves. \textit{The tea is bitter; it is aged in Aeler barrels, which add a taste of iron.}
\item[J] \textbf{The Bridge of Nine Echoes} — A rope bridge over a chasm; each step echoes nine times. \textit{The ninth echo is always a whisper; it is the name of the next person to fall.}
\item[Q] \textbf{The Viceroy’s Hall} — A longhouse of timber and stone; a throne of iron ore. \textit{The throne is cold; the viceroy never sits on it – only stands behind it.}
\item[K] \textbf{The Silent Pass} — A high mountain pass where no bird flies. \textit{The air is too thin for sound; arguments must be written in the snow.}
\item[A] \textbf{The Mine of the Ninth Seal} — A sealed tunnel with a Covenant seal on the timbers. \textit{Inside is a demon that the miners woke. The seal was placed by Brother Haridas. It is cracking.}
\end{enumerate}

\section*{Hearts — People \& Factions (Himadri Foothills)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Mine Foreman} — A Dhaharan with a dwarf’s whistle; he speaks both languages. \textit{He lost his hearing to a cave‑in; he reads lips in two scripts.}
\item \textbf{Elephant Mahout} — Old, missing two fingers, tattooed with elephant footprints. \textit{He knows the name of every elephant that has crossed the pass in fifty years.}
\item \textbf{Aeler Toll‑Keeper} — A dwarf with a brass scale and a ledger. \textit{He weighs your goods and your lies; the scale tips for both.}
\item \textbf{Raja’s Spymaster} — A woman with a limp; she uses a walking stick that is also a blowpipe. \textit{The darts are tipped with tea – caffeine that stops hearts.}
\item \textbf{Covenant Judge (Sister Gaura)} — Robed in grey, a water‑clock on her belt. \textit{The clock runs on tears; she refills it from her own eyes when she hands down a death sentence.}
\item \textbf{Smelter Worker} — A young man with red‑rimmed eyes. \textit{He coughs iron dust; his spittle stains the ground. He will not live past thirty.}
\item \textbf{Treasure Hunter} — A foreigner digging in an old landslide. \textit{He is looking for the tomb of a prince; the prince is the one who stole the amulet.}
\item \textbf{Aeler Engineer} — A dwarf with a tuning fork and a level. \textit{He can hear a crack in a beam from a hundred paces. He never walks under one.}
\item \textbf{Tea Trader} — A Fhara merchant with a caravan of mules. \textit{He sells tea to the Aeler; they pay in iron. He sells iron to the coast; they pay in silver.}
\item[J] \textbf{The Blind Pilgrim} — A woman in white, walking the passes alone. \textit{She is searching for a shrine where a bell rang without a clapper. The shrine is a cave.}
\item[Q] \textbf{The Ghost of the Landslide} — A figure in grey, seen only at dusk. \textit{It points at the pass. It is pointing at a safe route – or a trap.}
\item[K] \textbf{The Viceroy’s Cipher Clerk} — A dwarf who writes diplomatic letters in code. \textit{The code is a poem; the poem is about a lost mine. The mine is real.}
\item[A] \textbf{The Demon’s Whisper} — A voice heard in the sealed mine. \textit{It offers a deal: a year of good fortune for a single memory. The memory is the last breath of a dying friend.}
\end{enumerate}

\section*{Clubs — Complications/Threats (Himadri Foothills)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Cave‑In} — A tunnel collapses; the party is trapped. \textit{The collapse was not an accident – a saboteur loosened the timbers. The saboteur is in the party.}
\item \textbf{Stampede on the Pass} — Elephants panic; one throws its rider. \textit{The rider is the mahout; the elephant is mourning a dead calf. The calf was killed by a smuggler.}
\item \textbf{Toll Dispute} — The Aeler keeper demands double; the raja’s men refuse. \textit{The dispute is a trap; both sides want the party to choose a side. Neither is honest.}
\item \textbf{Landslide Ahead} — The pass is blocked; the alternative route is three days longer. \textit{The landslide was deliberate – a rival caravan master wants to delay the party.}
\item \textbf{Breath Shortage} — At high altitude, the party struggles. \textit{Mark Fatigue 1 per scene in high passes until they acclimatise. The acclimatisation requires a rest day.}
\item \textbf{Spy in Camp} — A worker is stealing iron samples. \textit{The spy is working for an Ashaani agent; the samples are for a new alloy. The alloy is a weapon.}
\item \textbf{Covenant Audit} — Sister Gaura demands to inspect the party’s cargo. \textit{The cargo includes a relic from the Red Desert – a stolen Covenant seal. The seal is a forgery.}
\item \textbf{Aeler Bell Code Change} — The toll‑keeper rings a different sequence; the party is now “unregistered”. \textit{The change was ordered by the viceroy; the viceroy is testing the party’s loyalty.}
\item \textbf{Elephant Sickness} — A disease spreads through the caravan. \textit{The disease is a curse from a local spirit; the spirit was disturbed by the mine’s expansion.}
\item[J] \textbf{The Ninth Echo} — On the Bridge of Nine Echoes, someone counts the echoes. \textit{The ninth echo says a name. The name is the speaker’s. They faint.}
\item[Q] \textbf{The Viceroy’s Summons} — Thrain demands the party attend him at his hall. \textit{He offers a deal: find the saboteur in exchange for a toll waiver. The saboteur is his own clerk.}
\item[K] \textbf{The Silent Pass’s Cold} — The temperature drops; the party loses 1 Supply per hour. \textit{The cold is supernatural – a frost spirit has taken residence. It wants a sacrifice.}
\item[A] \textbf{The Demon Wakes} — The sealed mine cracks open. \textit{A shadow escapes; it possesses the nearest Aeler engineer. The engineer begins to sabotage the pass.}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage (Himadri Foothills)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Iron Ingot} — A single bar of high‑grade iron; worth 1 Measure. \textit{The ingot is stamped with a dragon; the dragon is the viceroy’s mark.}
\item \textbf{Toll Waiver} — A brass token; pass any Aeler bridge for free. \textit{The token is a nail; it fits into a hole in the toll‑keeper’s scale.}
\item \textbf{Elephant Whistle (Himadri)} — A bone whistle; call an elephant from the caravan. \textit{The whistle is made from the tusk of a bull that died in a rockslide; its ghost runs with the elephant.}
\item \textbf{Aeler Breath‑Counter} — A silver bell; wear it to count your breaths in high altitude. \textit{The bell rings when you reach eight; stop, or the ninth will be your last.}
\item \textbf{Smelter’s Gloves} — Asbestos gloves; handle hot metal safely. \textit{The gloves are cursed; they itch when you lie.}
\item \textbf{Raja’s Signet Ring} — A silver ring with a tiger; command loyalty from the raja’s clan. \textit{The ring is a copy; the real one was lost in a landslide. The ghost of the original owner wears it.}
\item \textbf{Covenant Seal (Himadri)} — A clay seal; witness any oath in the foothills. \textit{The seal is a lump of clay; it must be heated to work. The heat comes from a coal taken from the sealed mine.}
\item \textbf{Map of the Silent Pass} — A chart drawn on leather; shows a safe route through the high pass. \textit{The map is incomplete; the missing section is a fold. The fold is a secret – a cave with a spring.}
\item \textbf{Viceroy’s Letter of Introduction} — A sealed scroll; grants access to any Aeler hold in the mountains. \textit{The letter is written in blood; the blood is the viceroy’s. He changes his blood type each season.}
\item[J] \textbf{The Blind Pilgrim’s Staff} — A wooden staff with a bell on top; ring it to summon a guide. \textit{The guide is a ghost – the pilgrim’s predecessor. She will lead you safely, but she will ask for a story.}
\item[Q] \textbf{The Demon’s Nail} — A piece of the sealed mine’s timber. \textit{Touch it to a lock; the lock opens. The lock also screams.}
\item[K] \textbf{The Ninth Echo’s Whisper} — A recording of the ninth echo; play it to silence a foe. \textit{The foe hears their own name and freezes for one exchange.}
\item[A] \textbf{The Viceroy’s Forgiveness} — A written pardon for any crime in the foothills. \textit{The pardon is conditional; the condition is a secret mission. The mission is to assassinate a rival viceroy.}
\end{enumerate}

\subsection*{Quick use notes (Himadri Foothills)}
\begin{itemize}
\item Draw until all four suits appear. Highest rank sets main Clock.
\item Diamonds are industrial and alpine tokens (ingots, whistles, seals) that shift Position.
\item If any A appears, echo \textbf{iron and breath} – mines that remember, bells that count, and passes that close behind you.
\end{itemize}

\subsection*{Himadri Foothills Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Mine:} Crumble a crust into the tunnel mouth. Ignore first \textit{Cave‑In} penalty.
\item \textbf{Knot the Bell Rope:} Tie a knot in the rope of the shrine bell. Cancel \textit{Ninth Echo} or take –1 die on your next stealth roll (the bell still rings, but muffled).
\item \textbf{Name to the Iron:} Whisper a dead miner’s name into an ingot. Gain +1 Effect when negotiating with Aeler this scene, but mark Obligation to the viceroy.
\end{itemize}

\begin{lorebox}
“The Himadri foothills are not a place; they are a negotiation. Every step is a toll, every breath a debt, every bridge a contract with a dwarf who remembers the face of your grandfather. Walk softly, pay in iron, and never – ever – count to nine.”

--- \textit{Dusana of the Raven Road, at the Bridge of Nine Echoes}
\end{lorebox}

\section{The Nine Kingdoms — ``The Peacock Throne and Its Rivals''}
\label{sec:nine-kingdoms}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere (Nine Kingdoms)}]
Between the Sarpies plateau and the Brass Coast lies a fertile, contested lowland: the land of the Nine Kingdoms. This is the heart of Dhahara, where ancient dynasties still claim descent from the sun, and where the Peacock Throne of Ahimsa casts its shadow over half a dozen rival courts. Here, rivers are borders, elephants are war machines, and every harvest is a negotiation between kings who have never met but whose grandfathers swore blood oaths. The Covenant has a strong presence in the sacred cities; the Aeler have counting houses in the iron towns; the Sidhi sail up the rivers to trade. But the Nine Kingdoms belong to the Dhaharan – to their poets, their temple dancers, their bandit kings, and their queens who rule from behind silk screens.

\vspace{2mm}
\textbf{What you notice first:} The smell of jasmine and dust. The way every crossroads has a painted stone – a boundary marker, a warning, a prayer. The constant murmur of bazaars where haggling is a form of theatre.

\textbf{The one rule that kills newcomers:} Never praise a king's elephant. The elephant is sacred; your praise will be taken as an insult to the king's own mount.
\end{tcolorbox}

\subsection*{The Nine Kingdoms (Core)}
The number “nine” is traditional; the actual number of kingdoms fluctuates with wars, marriages, and the mood of the Peacock Throne. The most influential are:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Kingdom} & \textbf{Capital} & \textbf{Ruling Clan} \\
\hline
Ahimsa & Peacock Throne City & Somavamsi (Sun Dynasty) \\
Vajrapur & Iron Gate & Kalachuri \\
Kalinga & Hidden Delta & Gangavamsi (Pirates) \\
Mahishmati & Buffalo Gate & Yadava \\
Anga & Anga Fort & Chandela \\
Surashtra & Western Port & Silhara \\
Kosala & Sravasti (Sacred City) & Pala \\
Magadha & Rajagriha & Gupta (former imperial) \\
Saurashtra (ninth) & Junagadh & Solanki \\
\hline
\end{tabular}
\end{center}

\subsection*{Prominent NPCs of the Nine Kingdoms}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Maharaja Suryavarman & King of Ahimsa (Peacock Throne) & Unite the nine under his dynasty & He has a pet peacock that follows him; the peacock’s cry is a warning of assassination attempts. \\
\hline
Rani Chandravati & Queen of Vajrapur & Protect the iron mines from Aeler encroachment & She wears a bracelet of dwarf teeth – each tooth is from an engineer she outwitted. \\
\hline
King Bhim of Mahishmati & Buffalo King & Expand his pastoral kingdom into Anga & He has never lost a buffalo race; his buffalo is named “Conqueror”. \\
\hline
Covenant High Judge Vasu & Head of the Kosala circuit & Enforce contract law across the kingdoms & He carries a scale that never balances; he says justice is never equal, only weighed. \\
\hline
Temple Dancer Mira & Devadasi of the Peacock Throne & Keep the peace through ritual & Her dance is said to bring rain; she has stopped three droughts. \\
\hline
\end{longtable}

\subsection*{Names of the Nine Kingdoms}
Sanskrit and Prakrit roots, with honourifics (Rao, Singh, Deva, Bai). Surnames often denote clan or profession.

\begin{longtable}{|p{0.2\textwidth}|p{0.2\textwidth}|p{0.25\textwidth}|p{0.25\textwidth}|}
\hline
\textbf{Male} & \textbf{Female} & \textbf{Clan/Surname} & \textbf{Epithet} \\
\hline
Surya & Chandravati & Somavamsi & the Peacock \\
\hline
Bhim & Mira & Kalachuri & Iron‑Fisted \\
\hline
Vasu & Ganga & Yadava & Buffalo‑Loyal \\
\hline
Rajendra & Lakshmi & Pala & the Learned \\
\hline
Krishna & Durga & Gupta & Coin‑Eater \\
\hline
\end{longtable}

\section*{Spades — Places (Nine Kingdoms)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peacock Throne Hall} — A vast chamber with pillars of carved marble; the throne is a peacock made of gold and lapis lazuli. \textit{The peacock’s eyes are rubies; they follow anyone who approaches the throne.}
\item \textbf{Iron Gate Market} — A bazaar outside Vajrapur’s fortress; the smell of iron dust and spices. \textit{The market has a section where slaves are sold; the slaves wear iron collars stamped with the king’s mark.}
\item \textbf{Buffalo Race Track} — A dusty oval outside Mahishmati; the betting is fierce. \textit{The track has a shrine at the centre; the shrine is to a buffalo that never lost. Its skeleton is on display.}
\item \textbf{Sacred City of Sravasti} — A walled city with a thousand temples; the Covenant has a courthouse here. \textit{The city has no bells; the Covenant forbids them. The silence is enforced by armed judges.}
\item \textbf{Anga Fort} — A hilltop fortress with a dry moat; the walls are painted with tiger stripes. \textit{The fortress has never been taken; the king’s secret is a network of hidden tunnels. One tunnel leads to the river.}
\item \textbf{Western Port of Surashtra} — A Sidhi trading post; the harbour is full of dhows. \textit{The port has a lighthouse shaped like a woman; the woman is a Sidhi saint. Her lamp is always green.}
\item \textbf{Rajagriha Ruins} — The former imperial capital, now a graveyard of broken pillars. \textit{The ruins are haunted by the ghosts of Gupta emperors. They ask travellers for news of their empire.}
\item \textbf{Elephant Stables of Ahimsa} — A complex of stone stalls; each elephant has its own name carved above the door. \textit{The elephants are fed betel nut to keep their tusks yellow. The stains are a sign of royalty.}
\item \textbf{Well of the Nine Kings} — A deep stepwell with nine landings; each landing has a statue of a different king. \textit{The water is black; it is said to show the face of the next king to die.}
\item[J] \textbf{The Silent Court of Kosala} — A Covenant courthouse with no speaking allowed; all arguments are written. \textit{The judges write their verdicts in disappearing ink; they are the only ones who know the secret reagent.}
\item[Q] \textbf{Peacock Gardens} — A walled garden behind the throne hall; only the king may enter. \textit{The garden contains a single peacock that is immortal; it was the first king’s familiar.}
\item[K] \textbf{The Whispering Pass} — A narrow defile between Anga and Mahishmati; the wind carries conversations. \textit{Spies use it to listen to enemy camps. The pass is named for the dead – they whisper too.}
\item[A] \textbf{The Nine Kings’ Tomb} — A subterranean complex where the founders of each dynasty are buried. \textit{The tomb is sealed; the key is a single word that changes each year. The word is the name of the current weakest king.}
\end{enumerate}

\section*{Hearts — People \& Factions (Nine Kingdoms)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peacock Throne Chamberlain} — A eunuch in silk; he controls access to the king. \textit{He has a mirror on a stick; he uses it to watch around corners. He has seen many plots.}
\item \textbf{Iron Mines Overseer} — A dwarf from Vajrapur; he wears an Aeler breath‑counter. \textit{He is loyal to the king, but his cousin is a viceroy’s clerk. He keeps two ledgers.}
\item \textbf{Buffalo Herder} — A young man with a staff and a red turban. \textit{He can find water in a drought by listening to the ground; his buffalo have never gone thirsty.}
\item \textbf{Covenant Judge (High Judge Vasu)} — A grey‑haired man with a brass scale. \textit{He has never smiled; his clerks have never seen him eat. He drinks only rainwater.}
\item \textbf{Temple Dancer (Mira)} — A woman with anklets of silver and eyes lined with kohl. \textit{Her dance is a prayer; the prayer is for peace. She has never failed – but the peace never lasts.}
\item \textbf{Bandit of Kalinga} — A man with a parrot on his shoulder, a scar across his face. \textit{The parrot is his conscience; it screams when he lies. He lies often.}
\item \textbf{Spice Merchant} — A Sidhi trader with a leather bag full of cinnamon. \textit{He speaks seven languages; he lies in all of them. His weights are true, though.}
\item \textbf{Poet of the Peacock Throne} — A woman with a stringed instrument, her voice cracked from singing. \textit{She writes verses for the king; the verses are encoded warnings. The king cannot read.}
\item \textbf{Well Keeper of the Nine Kings} — A blind woman who sits by the stepwell. \textit{She knows the depth of the water by the echo of a dropped stone; she knows the future by the same echo.}
\item[J] \textbf{The Ghost of Rajagriha} — A translucent figure in a crown. \textit{He asks one question: “Does my empire still stand?” The answer must be a lie, or he will weep and flood the ruins.}
\item[Q] \textbf{The Silent Judge’s Clerk} — A young woman who writes in disappearing ink. \textit{She is the only one who knows the reagent; she guards it with her life. She has been attacked twice.}
\item[K] \textbf{The Peacock Keeper} — A boy who tends the immortal peacock. \textit{The peacock speaks to him; it tells him which courtiers are plotting. The boy has never been wrong.}
\item[A] \textbf{The Ninth King’s Ghost} — A shade that walks the tomb. \textit{It is the ghost of the last king to be assassinated. It points to the killer – but the killer is always a living king.}
\end{enumerate}

\section*{Clubs — Complications/Threats (Nine Kingdoms)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Elephant Stampede (Royal)} — The king’s elephants break loose; they are trained to kill. \textit{The stampede was ordered by a rival queen; the target is a visiting diplomat. The diplomat is you.}
\item \textbf{Buffalo Race Riot} — A disputed finish; the crowd turns violent. \textit{The dispute is a trap; the losing king wants an excuse to declare war. Your presence is the pretext.}
\item \textbf{Covenant Excommunication} — High Judge Vasu declares your contract void; your goods are seized. \textit{The excommunication is a bribe; the bribe was paid by a merchant you undercut.}
\item \textbf{Temple Dance Cursed} — Mira stumbles; the rain does not come. \textit{The priest blames the party; the sacrifice is a goat. The goat is a spy.}
\item \textbf{Spy in the Peacock Gardens} — Someone overheard the king’s war plans. \textit{The spy is a poet; the poet is you – you were hired to write verses, not to eavesdrop.}
\item \textbf{Nine Kings’ Tomb Robbers} — Grave robbers break into the tomb. \textit{The robbers are agents of a rival kingdom; they seek the key word. The word is “forgiveness”.}
\item \textbf{Well of Nine Kings Poisoned} — The black water turns red; people sicken. \textit{The poison was placed by an Ashaani agent; the agent wants to destabilise the kingdoms.}
\item \textbf{Peacock Throne Assassination} — A courtier tries to kill the king. \textit{The assassin is the chamberlain; the chamberlain is a spy for Vajrapur.}
\item \textbf{Silent Court Leak} — A judge’s verdict is revealed before it is read. \textit{The leak is a clerk; the clerk is selling secrets to the Bandit King. The verdict is your death.}
\item[J] \textbf{The Whispering Pass Ambush} — The party is caught in a crossfire between two kingdoms. \textit{The pass is neutral ground; both sides will claim you are a spy. You must choose a side – or escape.}
\item[Q] \textbf{The Immortal Peacock’s Cry} — The peacock screams; the court panics. \textit{The cry is a warning; an assassin is in the hall. The assassin is a guest.}
\item[K] \textbf{The Ghost’s Demand} — The Ninth King’s ghost points at you. \textit{It accuses you of murdering the king’s descendant. The descendant is an infant. You must prove your innocence.}
\item[A] \textbf{The Well’s Vision} — One PC sees their own face in the black water. \textit{The face is old, scarred, and weeping. The well says: “This is your future if you stay.”}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage (Nine Kingdoms)}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peacock Feather} — A single iridescent feather; present it to enter any court. \textit{The feather is from the immortal peacock; it never fades.}
\item \textbf{Iron Gate Pass} — A brass key; bypass customs at Vajrapur’s market. \textit{The key is also a weapon; it can break chains.}
\item \textbf{Buffalo Horn Trumpet} — A carved horn; blow it to summon a buffalo cavalry. \textit{The cavalry are mercenaries; they demand payment in land.}
\item \textbf{Covenant Seal (Kosala)} — A clay seal; witness any contract in the sacred city. \textit{The seal is fragile; it breaks if the contract is false.}
\item \textbf{Temple Dancer’s Anklet (Nine)} — A silver anklet; wear it to gain +1 die to any persuasion roll. \textit{The bells are silent unless you are honest.}
\item \textbf{Rajagriha Coin} — A Gupta‑era gold coin; trade it for a favour from the ghosts. \textit{The ghosts will answer one question about the empire’s fall.}
\item \textbf{Well Water} — A vial of black water; drink it to see the future. \textit{The future is always a death – yours or someone else’s.}
\item \textbf{Silent Court Disappearing Ink} — A small bottle; write a message that vanishes. \textit{The reagent is your own tears; the message reappears when you weep.}
\item \textbf{Nine Kings’ Key Word} — A single word: “forgiveness”. Speak it to open the tomb. \textit{The word also opens a lock in the Peacock Throne Hall. The lock is a secret door.}
\item[J] \textbf{The Immortal Peacock’s Feather} — A second feather, from the tail. \textit{It is said to grant a wish; the wish is that the king will die.}
\item[Q] \textbf{The Whispering Pass’s Map} — A chart of the defile and its listening posts. \textit{The map is written on a human ear; the ear was from a spy.}
\item[K] \textbf{The Peacock Throne’s Blessing} — A royal decree; demand a favour from any king. \textit{The favour is limited: they will not start a war for you, but they will end one.}
\item[A] \textbf{The Ninth King’s Crown} — A gold circlet found in the tomb. \textit{Wear it to command the ghosts; they will fight for you once. Then they will claim you.}
\end{enumerate}

\subsection*{Quick use notes (Nine Kingdoms)}
\begin{itemize}
\item Draw until all four suits appear. Highest rank sets main Clock.
\item Diamonds are courtly and ritual tokens (feathers, seals, water) that shift Position.
\item If any A appears, echo \textbf{peacock and well} – feathers that grant audience, water that shows futures, and kings who play chess with peacocks.
\end{itemize}

\subsection*{Nine Kingdoms Superstitions (burn for +1 die once)}
\begin{itemize}
\item \textbf{Bread to the Peacock:} Crumble a crust before the throne. Ignore first \textit{Peacock Throne Assassination} penalty.
\item \textbf{Knot the Anklet:} Tie a silver bell to the dancer’s anklet. Cancel \textit{Temple Dance Cursed} or take –1 die on your next performance roll.
\item \textbf{Name to the Well:} Whisper a dead king’s name into the black water. Gain +1 Effect when negotiating a truce this scene, but mark Obligation to the Covenant.
\end{itemize}

\begin{lorebox}
“The nine kingdoms are not nine; they are a hundred, and they change with the monsoon. But the Peacock Throne is eternal – because it is empty. The king who sits on it is only a caretaker. The true ruler is the peacock, and the peacock does not speak.”

--- \textit{Dusana of the Raven Road, at the Peacock Throne Hall}
\end{lorebox}

\clearpage
\chapter*{Colophon}
\addcontentsline{toc}{chapter}{Colophon}

\vspace{2\baselineskip}
\begin{lorebox}
Dhahara is a ledger with two columns. The Mountain writes in slate; the Coast writes in ink. Neither column ever totals. The amulet Kalrantha is the missing line -- the error that neither side can account for. It was stolen by an Ashaani prince who carved his name in blood across the Way of Silk, and the desert he left behind is the interest that never stops compounding.

When you return it, you will not balance the ledger. You will only stop it from bleeding.

--- \textit{Dusana of the Raven Road, at Port Shed, during a monsoon that refused to end. The ink ran in rivulets.}
\end{lorebox}

\end{document}