\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{setspace}
\onehalfspacing

\title{The Dust Cycle\\\large Tales of the Uncounted — Stories from Below the Ledger}
\author{Dusana of the Raven Road}
\date{}

\begin{document}



\begin{quote}
\textit{``The rich write the ledgers. The poor live in the margins. And when the margins grow wide enough, the Hollow walks through them.''}\\
— Dusana of the Raven Road, at the edge of the Dye District
\end{quote}

\section*{Preface: The Weight That Does Not Count}
My people are the Tulkani. We raise no temples; we carry the hearth. Yet we also bear a memory: every door slammed in our faces, every cup left unfilled, every coin tallied and found wanting.

This winter I forsook the high roads and walked the alleys where the counting‑houses cast their shadows and the wealthy never tread. I listened to those who dwell in the cracks—beggars, guttersnipes, laundresses, bone‑pickers, costermongers, night‑soil men, and the women who scrub palace steps before dawn and vanish before the masters stir.

They gave me stories, not the tales of kings and heroes, but the ones the ledgers never record:

\begin{itemize}
\item The baker's daughter who taught herself to read from the rich’s discarded letters and then forged her own freedom.
\item The sewer‑rat who found a door that should not exist and stepped into a room filled with false silver.
\item The beggar king who bargained with the Hollow and still rules, because the Hollow has no use for a subject who gives it what it wants.
\item The laundress who washes blood from the garments of the powerful and keeps a ledger of her own, written in water, that cannot be burned.
\item The night‑soil man who bears the city’s waste on his back and knows every noble’s weak point, for he has seen their back doors at dawn.
\item The child who counted the steps from gutter to palace, found nine, took the ninth step, and became something else.
\item The moneylender’s son who, after a drunken killing of a rich patron, found a map to a hidden treasure and spent twenty‑years trying to atone—only to watch his riches grow with each act of contrition, until a beggar’s wink and a stray diamond reminded him that the Hollow laughs at all who try to buy forgiveness.
\end{itemize}

These are not stories of triumph; they are stories of survival—of those who learned that the ledgers do not count them and who learned to live in the spaces between the numbers.

The ink is ash and tears. The fire is borrowed. The road is still south. But this winter, the road is the alley.

\noindent— Dusana of the Raven Road, Silkstrand, in the hour before the ninth bell

\input{copyright}
\section{Part I: The Ledger's Shadow}
\subsection{What This Is}
\textbf{The Dust Cycle} is a collection of stories, rumors, and urban legends drawn from the lower classes of Fate’s Edge. It is not a mechanics‑heavy expansion—though it offers hooks, timers, and superstitions for GMs who wish to use them. It is foremost a resource for storytelling: a way to make the world feel lived‑in, to give voice to those usually silenced, and to remind players that the “heroes” are not the only people with stories worth telling.

The stories in this expansion are:

\begin{itemize}
\item \textbf{True, but not factual.} They are the kind of tales told in taverns and alleyways, passed mouth‑to‑mouth, changing with each telling. They may be exaggerations, misunderstandings, or outright inventions. Yet they are true in the way all good stories are true: they reveal something about the people who tell them.
\item \textbf{Salacious, scandalous, and sometimes cruel.} The poor are not saints; they gossip, scheme, and resent. The cycle reflects that honesty—showing pettiness, cruelty, and the capacity for both kindness and betrayal.
\item \textbf{Honest about the powerful.} The stories do not romanticize the poor nor demonize the rich. They acknowledge that the wealthy are also human—capable of generosity, stupidity, and genuine tragedy. The difference is that the poor have fewer choices.
\end{itemize}

\subsection{How to Use This Cycle}
Game Masters can employ this expansion to:

\begin{itemize}
\item \textbf{Add color and depth} to urban scenes. When the party walks a market or slum, they now have stories to hear.
\item \textbf{Seed adventures.} A rumor overheard in a tavern can become a heist, a rescue, or a mystery.
\item \textbf{Humanize the setting.} The poor are not mere background or victims; they are people with agency, wit, and their own agendas.
\item \textbf{Provide alternative perspectives.} The party may hear a version of events very different from the official record.
\item \textbf{Introduce the Ninth in everyday life.} The Hollow walks the alleyways too, and the poor have learned to listen for it.
\end{itemize}

\section{Part II: The Stories}
\subsection{The Storyteller's Note}
The following tales are told in the voices of the people who shared them with me. I have changed names and places to protect the living—and sometimes the dead. The stories are grouped by theme, but they are fluid; a laundress’s tale may also be a ghost story, a beggar’s tale may also involve a noble. They bleed into one another, like dust settling on everything.

\subsection{Chapter 1: The Beggar's Ledger}
\subsubsection{The King of the Gutter}
\textit{Told by Old Mikkel, who has sat at the same corner for forty years and has seen five mayors, three plagues, and one very confused demon.}

In every city there is a king who does not sit on a throne. He rules the alleys, the doorways, the places where the law does not go. In Silkstrand they call him the Beggar King. No one knows his true name—some say he was a noble who gambled away his fortune; others claim he is a Tulkani who never left. He wears a coat patched with a hundred fabrics and carries a staff forged from a broken oar.

He does not beg; he collects. Every beggar in the city pays him a tithe—a coin, a crust, a piece of gossip. In return he protects them: if a beggar is beaten, he knows the assailant; if one dies, he knows the burial site; if one disappears, he knows where to find them—or that they will not return.

The Watch tolerates him because he keeps the beggars from rioting. The nobles tolerate him because he keeps them from stealing. The beggars tolerate him because he is the only one who remembers their names.

One night a young thief asked the Beggar King: “Why do you do this? You could be rich. You could leave this place.”

The Beggar King laughed. “Leave? This is my kingdom. A king does not abandon his people. Besides,” he said, tapping his staff, “the Hollow walks the streets at night. Someone must remind it that the poor are still here.”

\begin{description}
\item[What's true] The Beggar King is a fixture of the Dye District, known for as long as anyone can remember. His network is extensive, and he knows everything that happens in the alleyways. His age is unknown—some say he has been there a century.
\item[What's exaggerated] He may not be a literal king, merely an old, well‑connected beggar. Yet the story is true in spirit: the poor forge their own hierarchies, often more organized than the rich imagine.
\item[What the poor believe] The Beggar King struck a bargain with the Hollow, giving it a name—his own or another’s—in exchange for the power to remember every face in his kingdom. He does not die because the Hollow does not forget him; it waits for him to err.
\end{description}

\textbf{Mechanical Hook:} The Beggar King can be contacted by anyone who knows the secret sign—a three‑tap knock on a doorframe. He will trade information for a story, a coin, or a favor. His network is vast, but he gives nothing for free. Lie to him, and he will remember.

\subsubsection{The Beggar's Court}
\textit{Told by Little Tilda, who is seven and has already learned that the best place to sleep is under the bakery's vent.}

In the back of the Dye District lies a courtyard absent from any map. The buildings lean so close they block the sky. The ground is packed earth, worn smooth by a thousand feet. At dusk the beggars gather there.

They call it the Beggar's Court—not a court, but a circle of the forgotten. The Beggar King sits on a crate at the center. The beggars bring him their grievances: a stolen blanket, a beaten child, a guard who took more than his due. He listens, then asks, “Who here can pay?”

If none can pay, the grievance is heard and noted; he will wait, patient. If someone can pay—a coin, a crust, a piece of gossip—he will act: sending a message, finding a witness, making the problem vanish.

The Watch knows of the Beggar's Court but does not enter. The Beggar King has a deal with the Watch: they leave his people be, and he does not encourage rebellion. The pact has held for forty years.

\begin{description}
\item[What's true] The Beggar's Court exists as a place where the poor air grievances and seek justice. The Beggar King is real, his network wide.
\item[What's exaggerated] He may not hold a literal court; he may simply be a trusted listener and helper. The story captures the spirit of the place.
\item[What the poor believe] The Beggar's Court is sacred ground; the Hollow cannot enter. The Beggar King made a pact with the stones themselves—he gave them a name, and they shield the court. Inside, you are safe; outside, you are on your own.
\end{description}

\textbf{Mechanical Hook:} The Beggar's Court is a safe haven. Anyone who enters is under the Beggar King's protection for one night, but the protection bears a cost: you must tell a story or leave a name. The Beggar King keeps a mental ledger of names; the Hollow respects it and does not visit him.

\subsection{Chapter 2: The Laundress's Lament}
\subsubsection{The Ledger in Water}
\textit{Told by Mother Yarrow, who washes clothes for three noble houses and has never lost a single stain—except one.}

A laundress sees everything: the blood on a lord's shirt after a duel, the wine on a lady's dress after an assignation, the ink on a merchant's cuffs after a night of “worrying about the books.” She does not judge; she washes and she remembers.

One laundress—Anya—served the same house for thirty years. She washed the master's shirts, the mistress's linens, the children's smallclothes. She watched the master’s children grow, the mistress age, the house fall into debt and rise again.

She kept a ledger—not on paper, which could be burned, but in her head, written in the stain patterns of the clothes she washed. A wine stain on a sleeve meant a secret meeting; a bloodstain on a cuff meant a duel; a particular fold in a shirt meant a certain servant had been beaten.

When Anya died, the mistress came to her funeral, weeping. “She knew everything,” she said. “And she never told a soul.”

The other servants nodded; they knew Anya had told some souls, and those souls were still telling.

\begin{description}
\item[What's true] Laundresses are often the best‑informed people in any household, discreet because speaking risks their livelihood.
\item[What's exaggerated] Anya’s “ledger in water” is likely metaphorical, but it is true that laundresses remember stains and the stories behind them.
\item[What the poor believe] Anya’s ghost still walks the laundry room, writing her ledger in the steam rising from the washbasins. Gazing into the steam reveals the house’s secrets—though it shows what you fear, not the truth.
\end{description}

\textbf{Mechanical Hook:} A character who works in a household for a season gains access to the household’s secrets. They may roll Wits+Insight (DV 3) to recall a useful piece of gossip. On a miss, the gossip is overheard by the wrong party. On a critical failure, the character gains Exposure +1.

\subsubsection{The Stain That Would Not Come Out}
\textit{Told by a young laundress who wished to remain unnamed—she was still employed by the house in question.}

Lady C favored a silk dress the colour of a winter sunset. She wore it to every ball, every dinner, every assignation. One night she did not return from a ball until dawn; her dress bore a dark stain. She said nothing, handed the dress to the laundress, and commanded, “Get this out. Do not ask where it came from.”

The laundress tried cold water, hot water, salt, lemon—nothing worked. The stain was not blood; it was something else, seeming to move when beheld.

She brought the dress to a hedge‑witch. The witch examined the stain and said, “This is not a stain. This is a name. Someone has written a name into the fabric. The name \textit{is} the stain.”

“Whose name?” the laundress asked.

“The name of someone who was not supposed to die,” the witch replied. “You cannot wash it out. You can only return the dress to the person who stained it.”

The laundress returned the dress to Lady C. Lady C screamed; the stain had spread, now covering the entire garment. Lady C burned the dress. That night her husband died in his sleep. The stain was his name.

\begin{description}
\item[What's true] Some stains cannot be removed; some names cannot be washed away.
\item[What's exaggerated] The stain may not have been literally magical—perhaps wine, ink, or something worse—but the story captures the idea that the poor know secrets the powerful cannot hide.
\item[What the poor believe] Certain stains are witnesses; they remember what happened. Trying to wash them away only makes them spread. The only way to remove a witness‑stain is to confess. Lady C did not confess; her husband’s ghost now haunts the laundry room of every house in the district.
\end{description}

\textbf{Mechanical Hook:} A character who discovers a witness‑stain may spend a scene studying it. On a successful Wits+Lore test (DV 4) they learn the name of the person who caused the stain and the crime. On a failure the stain spreads to the character’s own clothing, granting Exposure +1.

\subsection{Chapter 3: The Guttersnipe's Gospel}
\subsubsection{The Ninth Step}
\textit{Told by a child who answered only to “Girl.”}

Every alley contains a single step—the ninth step. It is neither first nor last. Step on it, and you will see what you ought not to see: a door that was not there, a face like yours but older, a warm coin.

The children know this. They count their steps—one, two, three, four, five, six, seven, eight—then pause, stepping over the ninth. The adults dismiss it as superstition, yet it is the adults who step on the ninth step and disappear.

One child—Pip—did not believe the tales. He stepped on the ninth step, saw a door, and walked through. He returned three days later wearing new shoes, unable to speak. He wrote in the dust: “The ninth step leads to a room. The room is full of coins. The coins are warm. The coins are not for us.” He has not spoken since; he counts his steps each time he walks.

\begin{description}
\item[What's true] Among Silkstrand’s children there is a belief that the ninth step is unlucky; they count to eight, then pause, and if they forget they go back and start again.
\item[What's exaggerated] The ninth step may not literally open a magic room, but the children are more attuned to the Ninth Taboo than many adults—because the poor have less to shield them from the Hollow.
\item[What the poor believe] The ninth step is a door through which the Hollow walks. To step on it is to invite the Hollow to follow. Pip stepped on it, and the Hollow followed him. His silence is not muteness; it is continued listening.
\end{description}

\textbf{Mechanical Hook:} In any city scene, a character who counts their steps and stops at eight gains +1 die on their next stealth roll (they have been careful). A character who steps on the ninth step must test Resolve (DV 4) or acquire the Marked by the Hollow condition for one day.

\subsubsection{The Coin That Spends Itself}
\textit{Told by a street urchin paid in bread to tell this story—and who ate the bread while telling it.}

There exists a coin that spends itself. It is neither silver nor gold; it bears the hue of a dead man's tooth. If you find it you can buy anything—a loaf of bread, a night’s shelter, a knife, a secret. But the coin bears a rule: you must spend it before sunset. If you do not, it will spend itself, buying something you did not wish to sell.

Mira found the coin in a gutter. She bought a hot meal, spent the coin, and was full. The next day she found another coin, bought a blanket, spent it, and was warm. The third day she bought a knife, spent the coin. She never found another coin. She searched every gutter, alley, and doorstep—nothing.

Then she found a dead beggar clutching a coin of the same colour and weight. She took it, bought a coffin, and buried the beggar. She never found another coin, yet she never went hungry again. The coin had spent itself, purchasing something she did not know she needed: a purpose.

\begin{description}
\item[What's true] Street children tell of “lucky coins”—coins that appear when most needed, often said to be the coins of the dead left to aid the living.
\item[What's exaggerated] The coin may not be literally magical, but it is true that the poor notice small gifts of luck—a dropped coin, a forgotten loaf, an unlocked door—more keenly than the rich.
\item[What the poor believe] The coin is a test. Spend it on yourself and it will leave you; spend it on another and it will stay. It is not a blessing but a lesson.
\end{description}

\textbf{Mechanical Hook:} Once per session a lower‑class character may declare they have found a “lucky coin.” They may spend it to gain +1 die on a single roll (the coin is spent on a bribe, a meal, or a distraction). If they do not spend it, the GM gains 1 SB. The coin watches.

\subsection{Chapter 4: The Night‑Soil Man's Truth}
\subsubsection{The Second Ledger}
\textit{Told by a night‑soil man who asked to be called “the Stink.”}

The night‑soil man carries the city’s waste on his back. He is invisible, despised, yet necessary. He also knows everything.

He sees the back doors of the great houses, knows which merchants receive late‑night visitors, which nobles are in debt, which servants are beaten and which are not. He keeps a ledger—not on paper, which could be burned, but in his head, written in the smells of the city. The scent of wine means a feast; blood means a murder; fear means a secret.

One night the Stink saw a man leaving a house at dawn, carrying a bleeding bundle. He followed the man to the river and watched him toss the bundle into the water, where it sank. The Stink told no one; who would believe a night‑soil man? But he remembered the smell, and he will always remember it.

\begin{description}
\item[What's true] Night‑soil men are among the most invisible and therefore most observant city‑dwellers.
\item[What's exaggerated] The Stink may not have literally witnessed a murder, but it is true that the poor know the secrets of the powerful.
\item[What the poor believe] The night‑soil man’s ledger is written in the air. If you can read it, you see the city’s truth. The Stink still walks the alleys, still remembers, and waits for the right person to hear him.
\end{description}

\textbf{Mechanical Hook:} A character who befriends a night‑soil man can gain access to his knowledge. On a successful Presence+Sway test (DV 4) the night‑soil man reveals a secret about a powerful person. On a failure he fears to speak, and the character gains Exposure +1.

\subsubsection{The Cart That Sings}
\textit{Told by a carter who was drunk and may have been lying—but his story was too good not to write down.}

There is a cart that sings. It is a night‑soil cart, one of those that carry the city’s waste, but it does not sing like a cart; it sings like a woman.

The carter who drives it—Souris—is deaf. He cannot hear the cart, but he can feel it: a vibration in his bones.

“The cart is sad,” he says. “It has carried the city’s waste for a hundred years. It knows all the city’s secrets. It is singing them to me.”

Other carters laugh at Souris, calling him mad, yet they refuse to ride with him at night, saying the cart’s song is too sad.

One night a young woman—Lina—rode with Souris. She heard the cart singing and wept.

“What did it sing?” Souris asked.

“It sang about a baby thrown into the river,” Lina replied. “It sang about a mother who never stopped looking.”

Lina never rode with Souris again, but she left a flower on the cart each morning. The cart’s singing grew quieter.

\begin{description}
\item[What's true] Many among the lower classes believe old objects can hold memories; a cart that has hauled waste for a century may have absorbed the city’s grief.
\item[What's exaggerated] The cart may not literally sing, but the poor are more prone to believe in the memory of objects than the rich.
\item[What the poor believe] The cart is not a cart; it is a witness. It has seen everything the city has thrown away. It still sings—for the dead, for the forgotten, for those never counted.
\end{description}

\textbf{Mechanical Hook:} A character who spends a night riding with the cart may roll Spirit+Lore (DV 4). On a success they learn a secret about the city’s past. On a failure they gain the Haunted condition for one day (they cannot stop thinking about the cart’s song).

\subsection{Chapter 5: The Dust‑Witch's Bargain}
\subsubsection{The Corner of the Eye}
\textit{Told by a dust‑witch who lived in a doorway and spoke in a voice that seemed to come from the cracks in the wall.}

There is a woman who lives in the dust. She is neither old nor young; she is the dust itself, present in every corner, crack, and space no one sweeps. She is the Dust‑Witch.

She does not ask for coin; she asks for stories. She listens, she remembers, and sometimes she acts.

A merchant lost a shipment of silk. He came to the Dust‑Witch and said, “Find my silk.”

The Dust‑Witch replied, “Tell me a story I have not heard.”

The merchant told a tale—about the day he cheated his partner out of a fortune.

The Dust‑Witch nodded. “Your silk is in the warehouse of your partner. He stole it. You stole from him. The silk is stolen. The truth is stolen. You are even.”

The merchant fumed, but he could not argue; the Dust‑Witch had spoken the truth.

\begin{description}
\item[What's true] The poor often turn to informal judges and advisors not recognized by the authorities—figures who may be cunning, wise, or simply lucky.
\item[What's exaggerated] The Dust‑Witch may not be literally magical, but she has learned to read people and to use their own stories against them.
\item[What the poor believe] The Dust‑Witch is older than the city. She was here before the first stone was laid; she will be here after the last stone crumbles. She is the city’s memory. Be honest with her, and she will be honest with you; lie, and she will find your lie in the dust.
\end{description}

\textbf{Mechanical Hook:} A character who seeks the Dust‑Witch must tell a true story. On a successful Presence+Performance test (DV 4) the Dust‑Witch provides useful information. On a failure she knows you have lied and will refuse to help you again.

\subsubsection{The Sweeper's Promise}
\textit{Told by a sweeper who had swept the same street for forty years and had never spoken to anyone—until that day.}

There is a sweeper who has swept the same street for forty years. He does not know why; he just sweeps—dust, leaves, scraps of paper.

One day a child asked him, “Why do you sweep the street?”

“Because I made a promise,” the sweeper replied.

“What promise?”

The sweeper did not answer; he kept sweeping.

The child watched him for a year. Every day the sweeper swept; every day the child watched.

One day the sweeper stopped, looked at the child, and said, “I promised my wife I would sweep this street until she came back. She died forty years ago. I am still sweeping.”

The child did not know what to say.

The sweeper smiled. “Do not be sad. I am not sweeping because I am waiting. I am sweeping because I am remembering.”

He kept sweeping.

\begin{description}
\item[What's true] Many perform small, repetitive acts of remembrance—sweeping streets, polishing steps, ringing bells—for the dead.
\item[What's exaggerated] The sweeper’s story may be metaphorical, but it is true that the poor create their own rituals of remembrance because they lack monuments.
\item[What the poor believe] The sweeper is not sweeping the street; he is sweeping the dust of the dead. The dust is the city’s memory. If no one swept it, the dust would rise and swallow everything.
\end{description}

\textbf{Mechanical Hook:} A character who helps the sweeper (by sweeping for a day or by telling him a story) gains +1 die on their next roll in the district. The sweeper’s promise is contagious—show kindness, and the city remembers.

\subsection{Chapter 6: The Costermonger's Heart}
\subsubsection{The Price of a Squab}
\textit{Told by a costermonger who sold pigeons and whose heart was as soft as the birds he sold.}

Finch sold pigeons in the market. He knew each bird by name, colour, and call.

One day a rich man approached his stall. “I want a pigeon,” he said.

“Which one?” Finch asked.

“The one you love most,” the rich man replied.

Finch looked at his birds. He knew which he loved most—a scruffy little bird with a broken wing, too old to sell, too ugly to eat. He kept it because it kept him company.

“I will not sell that one,” Finch said.

“I will pay you double.”

“No.”

“Triple.”

“No.”

The rich man laughed. “You are a fool. A bird is a bird.”

Finch met his gaze. “A bird is a bird. But a heart is a heart. And I will not sell mine.”

The rich man left. Finch kept the bird. It lived another year, then died. Finch wept.

The next day a young woman came to his stall. “I hear you are the one who will not sell his heart,” she said. “I need someone like that.”

They married a year later. The pigeon’s feathers were woven into her wedding veil.

\begin{description}
\item[What's true] The poor value loyalty; they cannot afford to buy things, so they treasure what they have.
\item[What's exaggerated] The story may be tidy, but it captures a real truth: the poor often possess a fierce sense of honor—because honor is the only thing they cannot be forced to sell.
\item[What the poor believe] Finch’s pigeon is still alive, somewhere. It is the bird that cannot die—because it was loved. Love something enough, and it will return to you.
\end{description}

\textbf{Mechanical Hook:} A character who shows kindness to an animal gains +1 die on their next social roll in the district (the costermongers will remember). A character who harms an animal gains Exposure +1—the market gossips.

\subsubsection{The Stolen Squab}
\textit{Told by a costermonger who was not Finch—and who was not entirely honest about his role in the story.}

Peck sold pigeons too, but he was no Finch. He was a cheat: he would take a squab from one stall and sell it at another, claim a thin bird was plump, charge double for a crooked‑legged bird.

One day Peck stole a squab from Finch and sold it to a rich man's cook.

The next day Finch’s stall was empty—all the birds were gone.

Peck felt no remorse; he was greedy.

That night he dreamed of thousands of birds pecking at his eyes. He woke blind.

He spent the rest of his life begging. He could not see, could not sell, could not cheat.

The rich man’s cook brought him a squab each day—the same squab he had stolen. The squab had grown old, too thin to eat, but it was the only food Peck could taste.

\begin{description}
\item[What's true] Cheats exist in every market; sometimes they are caught.
\item[What's exaggerated] Peck may not have literally gone blind, but the story captures the idea that the market has its own justice—often harsher than the law.
\item[What the poor believe] Peck is still alive, still blind, still eating the squab he stole. He has been eating it for fifty years. The squab does not die; it is the bird that cannot die—because it was stolen.
\end{description}

\textbf{Mechanical Hook:} A character who cheats a costermonger gains Exposure +1. The costermongers have long memories. If they remember you, the market will close to you. If they like you, they will give you discounts—and information.

\subsection{Chapter 7: The Washerwoman's Witness}
\subsubsection{The Door That Would Not Open}
\textit{Told by a washerwoman who worked at the river's edge and had seen a thousand secrets float past.}

Lira washed clothes at the river every morning. She washed the garments of the wealthy, saw the stains, knew the secrets.

One morning she saw a door. It had not been there the day before. It was a low, iron door set in the riverbank, crusted with rust.

Lira did not open the door. She had seen too many doors in her life; she knew some doors should not be opened.

But the door opened by itself. A pale, cold hand reached out, dropped a coin into Lira’s basket, then withdrew. The door closed—and vanished.

Lira told no one, but she kept the coin. It was warm. It was the colour of a dead man's tooth. It was the coin that spends itself.

She spent the coin on bread; she was full. The next day the door returned, opened again, and gave another coin. She spent it on coal; she was warm. The third day the door returned, opened again, and Lira did not take the coin. She closed the door herself.

The door did not return.

Lira told this story only once—she told it to me. Then she said, “I do not know what was on the other side of the door. But I know that it was waiting. It was waiting for me to open it. I did not. I hope that was the right choice.”

\begin{description}
\item[What's true] Washerwomen are often the first to notice strange things—the river carries secrets as well as laundry.
\item[What's exaggerated] The door may not have been literally magical, but the story captures the idea that the poor are often closest to the threshold.
\item[What the poor believe] The door still lies in the riverbank. It opens when the Hollow walks. If you see it, do not open it. But if you do, the Hollow will give you a coin—and that coin will buy you what you need…until it buys you.
\end{description}

\textbf{Mechanical Hook:} A character who encounters the door may choose to open it. On a success (Spirit+Resolve DV 5) they gain a coin that grants +1 die on their next roll. On a failure they gain the Marked by the Hollow condition. On a critical failure the door swallows them—they are lost for one scene.

\subsubsection{The River's Memory}
\textit{Told by a washerwoman who had washed clothes for sixty years and had seen everything—and who was still washing.}

The river remembers. It remembers every coin thrown into it, every body dropped, every secret whispered. It carries them all.

Mira could read the river. She would dip her hands into the water and feel the memories. She would know if a body had been thrown in the night before, if a coin had been lost, if a secret had been drowned.

One day Mira found a ring in the river. It was gold, engraved with a name—the same name as a woman who had vanished.

Mira did not keep the ring; she left it on the bank. The river took it back.

The next day the woman appeared. She was not dead, not alive; she was a ghost of the river.

“Thank you,” the woman said. “I was waiting for someone to find it.”

And then she was gone.

Mira never told anyone what the ring looked like or whose name was on it. She simply kept washing. The river remained full of memories.

\begin{description}
\item[What's true] The river does carry memories—bodies, coins, secrets. It is a witness.
\item[What's exaggerated] The ghost may not have been real, but the story captures the idea that the river is a place where the past lingers.
\item[What the poor believe] The river is alive. It is the city’s memory. If you throw something into it, the river will keep it for you—until you need it again.
\end{description}

\textbf{Mechanical Hook:} A character who searches the riverbank may find a clue. On a successful Wits+Investigation test (DV 4) they find a coin, a ring, or a piece of paper. On a failure the river takes something from them—a coin, a memory, a name.

\subsection{Chapter 8: The Moneylender's Map}
\textit{Told by Old Jorn, who now sits beside the Beggar King’s court, his hands trembling not from age but from the weight of a secret he has carried for twenty years.}

Jorn was the son of a moneylender who lived in the shadow of the Count’s manor. The Count was a man of polished manners and cruel appetites; he rode his chestnut stallion through the market each morning, scattering beggars like chaff before the wind. Jorn’s younger sister, Lina, loved to chase the stallion’s shadow, laughing as she darted between stalls. One fog‑laden dawn the Count’s horse reared, trampling Lina beneath its hooves. The Count did not stop; he merely tipped his hat and rode on, murmuring an apology that was swallowed by the crowd.

Jorn, then sixteen, watched his sister’s blood soak the cobbles. Grief curdled into a hot, bitter rage. That night, after drowning his sorrow in cheap ale, he stumbled into the Count’s manor through a servant’s entrance, found the Count slumped over a table, wine staining his lips, and drove a dagger into his throat. The Count gurgled once, then fell silent. Jorn snatched the Count’s purse, heavy with silver, and fled into the alleys.

Inside the purse he found a folded vellum, its edges singed from the Count’s candle. When he unfolded it by the gutter’s dim light, he saw a crude map—ink‑drawn, marked with a red X beside a stand of ancient oaks beyond the city walls. The map promised a treasure buried where the roots of the oldest tree met a stone shaped like a sleeping wolf.

Jorn kept the map hidden, fearing the Watch would hang him for murder if they learned of the Count’s death. For years he worked odd jobs—sweeping stables, hauling night‑soil, mending nets—saving every copper he could. When he was twenty‑four he finally felt brave enough to follow the map. He dug beneath the oaks, uncovered a stone wolf, and pried loose a chest of tarnished gold coins, jeweled goblets, and a single, flawless diamond the size of a pigeon’s egg.

The wealth should have freed him. Instead, each coin felt like a weight on his chest. He tried to assuage his guilt: he funded the orphanage that had taken in Lina’s friends, he paid for a widow’s medicine, he anonymously bought bread for the beggars at the Beggar King’s court. Every time he gave, the purse seemed to refill—more silver appeared in his pockets, more coins clinked in his satchel, as if the very act of giving summoned more riches from the Hollow’s ledger.

Twenty years after the murder, Jorn—now a gaunt man with a beard streaked gray—stood before the Beggar King’s court, his pockets heavy with newly minted coins. He approached a ragged beggar who sat shivering beneath a bakery vent, his eyes milky with age.

“Take this,” Jorn said, thrusting a pouch of silver into the beggar’s lap. “It is all I have. Let it buy you warmth, food, a bed.”

The beggar looked up, winked—a quick, knowing flicker of the lid—and shoved the pouch back. “Keep it, old man. The Hollow likes its debts paid in coin, not in kindness.” He rose, shuffled away, and vanished into the fog.

Jorn stared at the empty space where the beggar had been. His foot struck something hard on the cobbles—a glint of light caught his eye. He bent and picked up a diamond, identical to the one he had found in the treasure chest, warm to the touch, humming faintly as if it remembered the hand that had first lifted it.

A low, amused chuckle seemed to rise from the stones themselves, from the shadows between the buildings, from the very air. The Hollow laughed.

Jorn slipped the diamond into his palm, closed his fist, and walked away—not toward the counting‑houses, not toward the mansions, but deeper into the alleys where the poor still whispered their stories, where the ledger never counted them, and where the only true wealth was the memory of a sister’s laugh and the quiet certainty that some debts can never be paid with silver.

\begin{description}
\item[What's true] The story of a poor man who kills a noble in a drunken rage, finds a treasure map, and spends years trying to atone is a common whispered tale in the Dye District; many claim to have heard a similar account from a relative or a tavern regular.
\item[What's exaggerated] The map’s magical properties, the self‑refilling purse, and the returning diamond are embellishments that serve to illustrate the theme that guilt cannot be bought away.
\item[What the poor believe] The Hollow watches every attempt to bargain with it. When a man tries to buy forgiveness with riches, the Hollow merely laughs and returns the coin—sometimes as a diamond, sometimes as a memory. True atonement lies not in giving, but in remembering the cost of the taking.
\end{description}

\textbf{Mechanical Hook:} A character who discovers the Moneylender's Map may spend a day studying it. On a successful Wits+Lore test (DV 5) they learn the location of the treasure and gain a clue to its guardians. On a failure the map misleads them, leading to a dangerous encounter (gain Exposure +1). If the character attempts to use the treasure’s wealth to buy favors or absolution, they must test Resolve (DV 6); on a failure they gain the Condition Hollow’s Laughter (they suffer –1 die on all social rolls until they perform a genuine act of remembrance, such as telling the story of the victim to someone who never knew them).

\section{Part III: The Dust in the Ledger}
\subsection{Superstitions of the Uncounted}
The poor have their own relationship with the Ninth. It differs from the rich’s. The rich fear the Ninth because they have something to lose. The poor fear the Ninth because they have nothing else.

\begin{itemize}
\item \textbf{The Ninth Step (City):} In any alley the ninth step is a door. Step on it, and you will see what you ought not to see. The poor count their steps—eight, then a pause.
\item \textbf{The Ninth Coin:} A warm coin is a coin touched by the Hollow. Spend it quickly—or it will spend itself.
\item \textbf{The Ninth Breath:} When you enter a room, count the breaths of the others. If there are eight, you are safe. If there are nine, someone is not breathing—and they are listening.
\item \textbf{The Ninth Hour:} Midnight is the hour of the Hollow, but the poor know the Hollow is also awake at dawn—when night‑soil men work and the wealthy sleep.
\item \textbf{The Ninth Name:} Never speak the name of a dead person while you are eating. The dead might hear you. They might want to taste the food.
\item \textbf{The Ninth Stone:} Every threshold has a stone. If you find it and touch it, you will see who has crossed the threshold before you—and who has not come back.
\item \textbf{The Ninth Candle:} Light nine candles, and the Hollow will see you. Light eight, and it will pass you by.
\item \textbf{The Ninth Step (Ritual):} When you leave a place, walk eight steps and stop. Look back. The ninth step is for the Hollow.
\item \textbf{The Ninth Child:} A child who is the ninth‑born of a ninth‑born is a child of the Hollow. They will never die—but they will never truly live.
\item \textbf{The Ninth Door:} If you see a door that should not be there, do not open it. The door is not for you. It is for the Hollow.
\end{itemize}

\subsection{The Dust Cycle Timer}
The Dust Cycle Timer [6] tracks the accumulation of stories, secrets, and the weight of the poor. It ticks when:

\begin{itemize}
\item The party learns a story from the Dust Cycle (+1)
\item The party helps a poor person without expecting payment (+1)
\item The party harms a poor person for no reason (+2)
\item The party ignores a poor person's plea (+1)
\end{itemize}

When the timer fills, a story manifests. Roll on the Dust Cycle Manifestation table:

\begin{enumerate}
\item \textbf{The Beggar King's Favor:} The Beggar King sends a messenger. The party is offered a job—or a warning. If they refuse, the Beggar King becomes an enemy. If they accept, they gain a powerful ally.
\item \textbf{The Laundress's Secret:} A laundress approaches the party. She has a secret that could destroy a powerful person. She will trade it for protection.
\item \textbf{The Guttersnipe's Warning:} A street child warns the party of danger. The warning is cryptic, but it is true. The child will not explain further—they are afraid.
\item \textbf{The Night‑Soil Man's Truth:} A night‑soil man offers the party a truth. It is a secret about a powerful person. It is also a trap—if the party uses it, they will be blamed.
\item \textbf{The Costermonger's Gift:} A costermonger gives the party a gift—a squab, a loaf of bread, a flower. The gift is a token of trust. If they accept it, the costermonger will help them. If they refuse, they will be remembered.
\item \textbf{The Dust‑Witch's Bargain:} The Dust‑Witch appears. She offers the party a deal: a story for a story. If the party tells her a true story, she will tell them a secret. If they lie, she will curse them.
\item \textbf{The Moneylender's Map:} A grizzled figure offers a weather‑worn map that promises hidden riches. Accepting it grants a clue to a lost hoard, but using the wealth to buy favors summons the Hollow’s laughter (see Mechanical Hook above). Refusing the map earns the Beggar King’s respect (+1 to the Ledger of the Uncounted).
\end{enumerate}

\subsection{The Ledger of the Uncounted}
The Dust Cycle introduces a new player‑facing mechanic: \textbf{The Ledger of the Uncounted}. This shared resource tracks the party’s relationship with the lower classes.

\begin{description}
\item[The Ledger of the Uncounted] The party starts at 0. They gain +1 for each act of kindness to a poor person (not for payment—for genuine help). They lose -1 for each act of cruelty or indifference. The Ledger’s level affects social interaction in the city’s lower districts:
\begin{itemize}
\item \textbf{+3 or higher}: The poor actively help the party. They warn of danger, give discounts, and hide them from authorities. They gain +1 die on all social rolls in the lower districts.
\item \textbf{0 to +2}: The poor are neutral. They will not betray the party, but they will not help them either.
\item \textbf{-1 to -3}: The poor distrust the party. They will not warn of danger. They may actively report them to the authorities. They suffer -1 die on all social rolls in the lower districts.
\item \textbf{-4 or lower}: The poor are hostile. They will actively work against the party—setting traps, spreading rumors, alerting the Watch. The party is no longer safe in any alley or doorway.
\end{itemize}
\end{description}

The Ledger can be raised by genuine acts of kindness—sharing food, protecting a child, telling a story. It can be lowered by cruelty, indifference, or simply ignoring the poor.

\section{Part IV: The Dust Cycle Generator}
\subsection{The Storyteller's Note}
The following generator tables are designed to create Dust Cycle stories on the fly. Use them to seed rumors, encounters, and adventures in the lower districts of any city in Fate’s Edge.

\subsubsection{Spades: The Place}
Roll or choose.

\begin{description}
\item[2.] \textbf{The Alley of the Ninth Step}—An alley where the ninth step is a door. Locals avoid it; strangers are warned. Some do not listen.
\item[3.] \textbf{The Beggar's Court}—A courtyard where the poor gather. The Beggar King holds court here. The Watch does not enter.
\item[4.] \textbf{The Laundry of Secrets}—A washhouse where the poor wash the clothes of the rich. The stains have stories; the stories are sold.
\item[5.] \textbf{The Night‑Soil Man's Route}—A path through the city that night‑soil men take. It is the most direct route—and the most hidden.
\item[6.] \textbf{The Dust‑Witch's Corner}—A doorway where the Dust‑Witch lives. It is always just out of sight. If you find it, you have been looking.
\item[7.] \textbf{The Costermonger's Stall}—A market stall where birds are sold. The costermonger knows everything. He listens to the birds.
\item[8.] \textbf{The Washerwoman's Riverbank}—A place by the river where women wash clothes. The river is a witness. It remembers.
\item[9.] \textbf{The Sweeper's Street}—A street that is always clean. The sweeper never stops. He is sweeping for the dead.
\item[10.] \textbf{The Gutter's Mouth}—A drainpipe that leads to the city’s bowels. It is a passage for the desperate and the brave. It leads to a place not on any map.
\item[J.] \textbf{The Uncounted Door}—A door that appears only at dawn. It leads to a place where the poor go to trade secrets. The door is not always there; it is always remembered.
\item[Q.] \textbf{The Ninth Threshold}—A doorway that should not be there. It is the Hollow’s door. The poor do not go near it—unless they have nothing left to lose.
\item[K.] \textbf{The Ledger's Shadow}—A counting‑house that is not on any map. It is where the poor keep their own ledgers. The ledgers are written in tears. They are read by the Hollow.
\item[A.] \textbf{The Dust's Heart}—A place where the dust of the city gathers. It is the city’s memory. The Dust‑Witch lives here. She is older than the city.
\end{description}

\subsubsection{Hearts: The Actor}
Roll or choose.

\begin{description}
\item[2.] \textbf{The Beggar King}—The ruler of the alleyways. He knows everything. He charges a price.
\item[3.] \textbf{The Laundress}—A woman who washes clothes and remembers stains. She is the city’s memory.
\item[4.] \textbf{The Night‑Soil Man}—A man who carries the city’s waste. He is invisible. He sees everything.
\item[5.] \textbf{The Guttersnipe}—A child who lives on the streets. They have seen the door. They have not stepped through—yet.
\item[6.] \textbf{The Costermonger}—A man who sells birds. He knows the market. He knows the people. He knows the secrets.
\item[7.] \textbf{The Washerwoman}—A woman who washes at the river. She has seen a thousand secrets float past.
\item[8.] \textbf{The Sweeper}—A man who sweeps the same street every day. He is waiting. He will not say for what.
\item[9.] \textbf{The Dust‑Witch}—A woman who lives in the dust. She knows all. She charges a story.
\item[10.] \textbf{The Miller's Daughter}—A woman who lost everything. She lives in the gutter. She remembers being rich. She does not speak of it.
\item[J.] \textbf{The Uncounted Child}—A child who was not counted in the census. They exist outside the ledgers. The Hollow does not see them—yet.
\item[Q.] \textbf{The Ninth‑Born}—A child who is the ninth‑born of a ninth‑born. They are touched by the Hollow. They will never die—but they will never truly live.
\item[K.] \textbf{The Ledger‑Keeper}—A poor man who keeps his own ledger. He counts the debts of the rich. He is waiting for the day they are called in.
\item[A.] \textbf{The Dust's Daughter}—A woman who lives in the dust. She is the Dust‑Witch’s apprentice. She will become the Dust‑Witch when the old one dies. She is not sure she wants to.
\end{description}

\subsubsection{Clubs: The Pressure}
Roll or choose.

\begin{description}
\item[2.] \textbf{A Noble's Cruelty}—A wealthy person is harming the poor. The victim is someone the party knows—or should protect.
\item[3.] \textbf{A Rival's Scheme}—Another thief or criminal is claiming territory. The poor are caught in the middle.
\item[4.] \textbf{A Watch Raid}—The Watch is cracking down on the poor. They are looking for a specific person—or someone to blame.
\item[5.] \textbf{The Hollow's Curiosity}—The Hollow has noticed the poor. It is asking questions. The poor are afraid to answer.
\item[6.] \textbf{A Lost Coin}—A coin from the Dust Cycle has appeared. It is warm. It is spending itself. Someone is going to lose something.
\item[7.] \textbf{A Stolen Secret}—A secret from the poor’s ledger has been stolen. It is being used against them. They cannot afford to fight back.
\item[8.] \textbf{A Broken Promise}—Someone made a promise to the poor. They broke it. The poor are angry. They are taking revenge.
\item[9.] \textbf{A False Witness}—Someone is lying about the poor. They are being blamed for a crime they did not commit. The real culprit is a powerful person.
\item[10.] \textbf{A Blood Debt}—Someone in the party owes the poor a debt. The debt is being called in. It cannot be refused.
\item[J.] \textbf{The Ninth Step}—Someone has stepped on the Ninth Step. The Hollow is following them. The poor are afraid—it will lead the Hollow to them.
\item[Q.] \textbf{The Uncounted Door}—The Uncounted Door has appeared. It is open. Something has come through.
\item[K.] \textbf{The Dust's Anger}—The Dust‑Witch is angry. Someone has lied to her. She is cursing the culprit. The curse will spread to everyone in the district.
\item[A.] \textbf{The Ledger's Reckoning}—The poor are demanding justice. They are holding a tribunal. The accused is a powerful person. The judge is the Beggar King. The sentence will be harsh.
\end{description}

\subsubsection{Diamonds: The Leverage}
Roll or choose.

\begin{description}
\item[2.] \textbf{The Beggar King's Favor}—The Beggar King owes the party a favor. He will help them once. The favor will be granted—but it will be expensive.
\item[3.] \textbf{The Laundress's Secret}—A laundress knows a secret that could destroy a powerful person. She will trade it for protection.
\item[4.] \textbf{The Night‑Soil Man's Route}—A night‑soil man gives the party a route through the city. It is secret. It is safe. It is not on any map.
\item[5.] \textbf{The Guttersnipe's Warning}—A street child warns the party of danger. The warning is cryptic. It is also true.
\item[6.] \textbf{The Costermonger's Gift}—A costermonger gives the party a gift. It is a small thing—a squab, a loaf of bread, a flower. It is a token of trust.
\item[7.] \textbf{The Washerwoman's Witness}—A washerwoman offers to testify for the party. She saw something. She will speak. She is not afraid.
\item[8.] \textbf{The Sweeper's Promise}—A sweeper offers to help the party. He knows the street. He knows the people. He will open doors—if he is asked.
\item[9.] \textbf{The Dust‑Witch's Bargain}—The Dust‑Witch offers the party a deal. A story for a story. She will trade a secret for a secret.
\item[10.] \textbf{The Uncounted Child's Offering}—A child offers the party a coin. It is warm. It is the coin that spends itself. It is a test.
\item[J.] \textbf{The Ninth‑Born's Vision}—A child touched by the Hollow sees something. They will tell the party—if they ask nicely.
\item[Q.] \textbf{The Ledger‑Keeper's Ledger}—A poor man shows the party his ledger. It records the debts of the rich. It is a weapon.
\item[K.] \textbf{The Dust's Daughter's Warning}—The Dust‑Witch’s apprentice warns the party. The Dust‑Witch is angry. Someone is going to be cursed.
\item[A.] \textbf{The Dust's Heart}—The party finds the Dust’s Heart. It is the city’s memory. They can read it. The cost is a memory of their own.
\end{description}

\section{Part V: The Dust Cycle Manifestations}
\subsection{The City's Memory}
The Dust Cycle is not just a collection of stories—it is the city’s memory. The poor remember what the rich have forgotten. They remember the names of the dead. They remember the secrets of the powerful. They remember every injustice.

If a character spends a scene in the Dust’s Heart (the place where the city’s memory gathers), they may attempt to read the memory of the city. On a successful Spirit+Lore test (DV 5), they learn a secret about the city’s past. On a failure, they gain a memory that is not their own. On a critical failure, they lose a memory of their own.

\subsection{The Dust‑Witch's Curse}
The Dust‑Witch is not evil. But she is old. She is patient. And she is tired of being forgotten.

If a character lies to the Dust‑Witch, she will curse them. The curse is not dramatic—it is subtle. It is the curse of being forgotten. The character will find that doors close in their face, that people do not remember their name, that their coins are always a little light.

The curse can be lifted by telling the Dust‑Witch a true story—a story that she has not heard before. The story must be worth the curse.

\subsection{The Beggar King's Ledger}
The Beggar King keeps a ledger in his head. It is the record of every debt, every promise, every secret in the alleyways. He does not write it down—he remembers.

If the party does a great favor for the Beggar King, he will add them to his ledger. This means they are known. They are watched. They are protected. It also means they are owed.

If they betray him, he will erase them from his ledger. This means they are forgotten. They are not protected. They are not owed. They are nothing.

\section*{Epilogue: The Ledger of the Uncounted}
I closed the notebook. The fire was low. The alley was empty.

I had written down the stories of the poor. I had listened to their voices. I had done my best to tell them true.

But I knew I had not told all the stories. There were always more. The city was full of them. They were in the dust, in the stains, in the things that no one swept away.

The poor do not need heroes. They need witnesses. They need someone to remember their stories.

I am a witness. I have remembered.

And now, these stories are yours.

The ink is dust. The fire is cold. The road is south. The poor are still counting.

— Dusana of the Raven Road, at the edge of the Dye District, in the hour before the ninth bell

\section*{Colophon}
This book was written on paper made from rags, in ink cut with ash and the tears of a washerwoman. The binding is leather from a costermonger's donkey.

The ninth story is the one I did not tell. It is the story of the Hollow's own beginning. It is not a story that belongs to the poor. It is a story that belongs to the Hollow itself.

The poor do not speak of it. They do not need to. They already know what it is like to be forgotten.

— Dusana of the Raven Road
\end{document}
