\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{quoting}

ℎypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Lethai Module for Fate's Edge},
    pdfauthor={AI Assistant}
}

\setlist{leftmargin=*}
\setlist[itemize]{itemsep=0pt}
\setlist[enumerate]{itemsep=0pt}

≠wcommand{\talent}[1]{\textbf{#1}}
≠wcommand{\stringtype}[1]{\textbf{#1}}
≠wcommand{⨾licationtype}[1]{\textbf{#1}}
≠wcommand{\gift}[1]{\textbf{#1}}

\title{Lethai Module for Fate's Edge}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Core Concept}
The Lethai are divided by an ancient curse that prevents them from fully accessing both physical and mental gifts. This division creates distinct elven cultures, each excelling in different aspects of existence while yearning for wholeness. They are masters of living law, environmental stewardship, and contextual knowledge.

\section*{Affinity (All Lethai)}
\textbf{Gift of the Body OR Gift of the Mind (Choose One):}
\begin{itemize}
  ℹtem \textbf{Gift of the Body (Lethai−al/ar):} +1 die to physical actions (Body−based skills). Once per scene, can spend 1 Boon to enhance a physical ability beyond normal limits for one action.
  ℹtem \textbf{Gift of the Mind (Lethai−thora):} +1 die to mental actions (Wits/Spirit−based skills). Once per scene, can spend 1 Boon to recall or deduce information that would normally be beyond immediate knowledge.
\end{itemize}

\textbf{The Curse of Division:} Lethai cannot take talents that enhance both physical and mental capabilities simultaneously. If a talent would grant bonuses to both, choose one benefit.

\section{Lethai−al −−− Root, River, and the Roof−Tree}

\subsection*{Woodwise lawkeepers, merchants of light and shade, and stewards of living borders}

\begin{quote}
``Name yourself once, name the river twice, and never name the forest as if it were yours.''
\end{quote}

\medskipℎrule\medskip

\subsection{I. Ground & Memory}

The Lethai−al dwell where roof−trees braid the sky and rivers think aloud. Their memory is arboreal: rings and seasons, storms and healings, disputes recorded in coppice and replanting rather than tablets and seals. A Lethai−al oath is \textbf{root−law}: it binds living things−−−people, paths, waters−−−and is paid in years. To outsiders they appear quiet; to neighbors they are relentless auditors of footprint and flow.

\medskipℎrule\medskip

\subsection{II. Forest Commons (Places that Hold)}

⫕section*{Valewood}
Tall−canopy forest stitched with rain gardens and stone−steps. Trails shift by season; waystones carry moss−tallies (green for open, gray for HEAL, black for trespass). Songs mark crossings where words would bruise the undergrowth.

⫕section*{Roof−Tree Circles}
Villages built around elders: one monumental tree per circle, scaffolded with walkways and rain−catchers. Roof−Tree sap is consecrated−−−used to seal pacts and to dismiss them when harms outweigh promises.

⫕section*{River Courts}
Every significant stream keeps a court: a shingle bank, a willow−bench, a book of pebbles. Trade, ferry rights, and flood work are judged where the water can overhear.

⫕section*{Edge−Wards}
Where wood meets field or steppe, boundary groves are tended as treaties: thorn, ash, and fruit interplanted. The sharp feeds the soft−−−hedge law that stops hooves and nourishes neighbors.

\medskipℎrule\medskip

\subsection{III. Courtesy & Everyday Law}

\textbf{Shade Etiquette} is the grammar of Lethai−al life. Observing it grants hospitality and hearing; ignoring it turns the forest against your plans.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Iron Covered.} Bare iron offends ward−lines; wrap it in leather or cloth. (Mechanic: on first entry, covered iron grants \emph{Position +1} in any parley or request to pass.)
  ℹtem \textbf{Name Once.} Speak your name and intent at the edge; do not name the forest as property.
  ℹtem \textbf{Step on Stone.} Where stones are laid, use them; every crushed sprout is a debt.
  ℹtem \textbf{Water First.} Pour a first cup to the river or cistern before you drink.
  ℹtem \textbf{Leave the Light.} Replace the shade you take: plant, HEAL, or pay. (Use \emph{Light−Due Strings} below.)
\end{itemize}

\textit{Timers for breaches:} \textbf{Under−Root Grudge [4]}, \textbf{Stream−Clouding [4]}. When a timer fills, expect formal censure, raised tariffs, or wardens on the path.

\medskipℎrule\medskip

\subsection{IV. Courts & Factions (Not One Voice)}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Wardens of the Roof−Tree.} Rangers and surveyors who measure canopy health, poacher pressure, and firebreaks. They carry leaf−badges and thorn−writs.
  ℹtem \textbf{Merchant Foresters (Syndics of Light).} Manage sustainable fellings, resin trade, boat−timber quotas, and lantern−wood auctions. They argue that trade funds stewardship.
  ℹtem \textbf{Songkeepers.} Priests of repair rites and remembrance festivals. They arbitrate forgiveness by counting regrowth, not coin.
  ℹtem \textbf{Ferrymen of the Rain.} River pilots and dam−tenders; their votes swing courts in flood years.
\end{itemize}

\textit{Tension lines:} quota vs. canopy; export timbers vs. local craft; river straightening for mills vs. flood meadows for salmon.

\medskipℎrule\medskip

\subsection{V. Borders & Neighbors}

⫕section*{Silence Furlong (with the Ykrul)}
A speechless border strip: no grazing, no felling, no names. Cross in silence, then speak once. Honored, it keeps peace warm; broken, gray−fletched messengers appear at dusk. Joint patrols HEAL hedges and lift stones; concessions are marked as seasonal strings on both boards.

⫕section*{Ridge Courtesy (with the Aeler)}
Root−law and stone−law trade proofs. Aeler build keystone stairs where Lethai−al choose the line; Lethai−al plant edge−wards that keep roads alive. Disputes end with bread in oven courts and sap at roof−trees.

⫕section*{Moots at the Mouth (with the Linns)}
Ship−timber quotas bind with replanting moots and harbor dues. When storm seasons bite, both sides sing the river clear.

\medskipℎrule\medskip

\subsection{VI. Economy of Light (Strings & Ledgers)}

Light and shade are currencies. Every fell, ferry, and fire has a \textbf{light−due}−−−a balance to be repaid in replanting, canal−clearing, or trade at kinder rates.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Strings (examples):} \emph{light−due receipt}; \emph{shade−credit}; \emph{ferry right}; \emph{resin share}; \emph{canoe−lane priority}; \emph{seed tithe}.
  ℹtem \textbf{Use:} Cash a light−due to gain \emph{DV −−1} on any operation framed as repair, replanting, or flood work. Abuse it and start \emph{Under−Root Grudge [1]}.
  ℹtem \textbf{Quotas as Timers:} \emph{Boat−Timber Quota [6]}, \emph{Lanternwood Allotment [4]}. Fill the timer to unlock export; overfill and trigger \emph{Canopy Censure}.
\end{itemize}

\medskipℎrule\medskip

\subsection{VII. Threats & Remedies}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Bark Blight.} A fungal rot spread by careless carts and damp blades. Remedy: ash−wash stations, tool quarantine, fire−lines by moon.
  ℹtem \textbf{Poacher Pressure.} Syndicates take ripe resin and seedwood without dues. Remedy: undercover ferries; tariff traps at market; public shaming rites.
  ℹtem \textbf{River Clouding.} Silt from mills and bad embankments chokes fish beds. Remedy: settle−ponds, scheduled sluices, mill−weirs redesigned with Aeler math.
  ℹtem \textbf{Old Fire.} Lightning scars that wake in drought years. Remedy: fuel mosaics, controlled burns, ember patrols.
\end{itemize}

\textit{Fronts (example):} \textbf{Bark Blight [6]} → \textbf{Canopy Hunger [4]} → \textbf{Famine Tariff [4]}.

\medskipℎrule\medskip

\subsection{VIII. Tools & Tokens (at the table)}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Leaf−Badge (Warden).} Once/scene, upgrade Position by +1 when enforcing forest law or rescue.
  ℹtem \textbf{Seed−Tithe Seal.} Convert a sanction into a \emph{Repair Project [4]} instead of a fine or exile.
  ℹtem \textbf{Ferry−Knot.} Skip the first \emph{River Hazard} consequence on a leg or chase; on use, pay a \emph{light−due}.
  ℹtem \textbf{Moss−Tally.} Reveal one hidden quota timer in a market or lumberyard scene.
\end{itemize}

\medskipℎrule\medskip

\subsection{IX. Aesthetics & Speech}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Materials:} bark−laminate armor, willow−bone frames, resin−glass panes, river−iron sheathed in leather.
  ℹtem \textbf{Colors:} rain green, bark gray, amber, sky−through−leaf.
  ℹtem \textbf{Proverbs:}
    \begin{itemize}
      ℹtem ``Leave the forest no thinner than your shadow.''
      ℹtem ``The river remembers who spoke over it and what they promised.''
      ℹtem ``Cut where it wants to fall or do not cut at all.''
      ℹtem ``If the hedge is hungry, the treaty is thin.''
    \end{itemize}
  ℹtem \textbf{Names:} given + grove + season/deed: \emph{Talan of Reedfall, Who Unknotted the Flood}.
\end{itemize}

\medskipℎrule\medskip

\subsection{X. Integration Hooks}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Political Intrigue:} Forest syndics lobby city councils for lanternwood allotments; bribes look like "donations" to flood work.
  ℹtem \textbf{Caravans:} Resin and seed routes require \emph{shade−credit}; escorts must mind iron covers and Silence Furlongs.
  ℹtem \textbf{Wilderness:} Use \emph{Hinterlands} procedures with \emph{fuel mosaics} and \emph{blight quarantine} as special orders.
  ℹtem \textbf{Violets & Stone:} Dock courts host River Days; Lethai−al ferries grant \emph{canoe−lane priority} to crews who HEAL ladders and steps.
\end{itemize}

\medskipℎrule\medskip

\subsection{XI. Adventure Seeds}

\begin{enumerate}[leftmargin=*]
  ℹtem \textbf{The Hedge That Ate a Road.} An edge−ward grew too well and swallowed a toll lane. Broker a cut and a replanting that keeps both law and trade.
  ℹtem \textbf{Blight Under Boots.} Bark Blight rides a caravan's wheel hubs. Build ash−wash stations without sparking a tariff war.
  ℹtem \textbf{Silence Furlong, Twice Broken.} Two violations in one season risk a border feud. Walk the strip, plant the apology, and write a geometry both sides can live in.
  ℹtem \textbf{River Court at Flood.} Mills want straighter banks; fishers want flood meadows; Aeler offer weirs. Choose a design, then stand in the water and defend it.
  ℹtem \textbf{Lanternwood Ledger.} A stenographer vanished with the quota book before an export moot. Find them, or reconstruct the truth from moss tallies and ferry knots.
\end{enumerate}

\medskipℎrule\medskip

\subsection{XII. Portraying the Lethai−al}

Let \emph{policy} speak before mystique: quotas, ferries, hedges, and the cost of light. Make courtesy concrete (stones, coverings, cups) and consequences seasonal, not theatrical. Honor the forest as a \emph{party to the treaty}, not scenery. When in doubt, ask: \emph{What did this footstep cost, and who will we pay?}

\medskipℎrule\medskip

\subsection{XIII. Language & Time (Context is King)}

Lethai speech is \textbf{context−saturated}. Syntax legible to grandparents will baffle grandchildren unless framed by the right place, season, and kinship. Meaning travels with \emph{context keys}−−−gesture, setting, shared story. A sentence lifted from its grove can die on the road.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Context Keys.} Place−name, season−mark, kinship−hand, and roof−tree sign. Absent any two, assume misreadings.
  ℹtem \textbf{Old Texts.} Manuscripts older than two generations require \emph{gloss−trees} (marginal twig glyphs) or a \emph{songkeeper} to render.
  ℹtem \textbf{In play.} When reading or pleading in older registers, increase DV by +1 unless the party holds a \emph{Context String} (gloss−tree, witness, song). Cashing a Context String reduces DV by −−1 and creates a \emph{Shared Frame} for the scene (Position +1 on first exchange).
\end{itemize}

\medskipℎrule\medskip

\subsection{XIV. Peoples & Lineages (Not One Grove)}

\paragraph{Lethai−al −−− \emph{The People of the Body}} 
The woodwise many. Found in most great forests, they are the elves most humans meet and the most willing to integrate in mixed cities. \textbf{Embodied and present}: heightened senses, quick strength, an ease with weather and work. Most \emph{half−elves} trace lineage here.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Everyday.} Ferry crews, hedge−wardens, resin syndics, ward dancers.
  ℹtem \textbf{Gifts (examples).} \emph{Canopy Spring} (Position +1 on leaps/climbs once/scene); \emph{Scent of Rain} (clue on approaching weather/rot); \emph{Root−Balance} (resist shove/knockdown).
  ℹtem \textbf{Strings.} ferry knot; resin share; roof−tree blessing.
\end{itemize}

\paragraph{Lethai−ar −−− \emph{The Oathbound (Dark Elves)}} 
Wood elves who swear to \textbf{Inae} (the Weaver, Angel of Spiders) or \textbf{Isoka} (the Serpent Queen, Angel of Snakes). Rare in the upper Amaranthine, growing along shaded trade arteries and cavernous underways. Their vows shape craft and law: webs of obligation, clean strikes, poison as \emph{medicine with teeth}.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Everyday.} Understory judges, silk engineers, antidote keepers, night pilots.
  ℹtem \textbf{Gifts (examples).} \emph{Weaver's Reading} (trace networks: DV −−1 on uncovering plots/paths); \emph{Stillness} (be unnoticed in dim light until you move to act); \emph{Venin Lore} (convert Harm 1 from toxins to Fatigue once/scene).
  ℹtem \textbf{Strings.} spider−vow token; serpent oath; silk−tithe.
  ℹtem \textbf{Courtesy.} Oath−signs worn openly; breaking web−law costs \emph{Mask Rights} for a season.
\end{itemize}

\paragraph{Lethai−thora −−− \emph{The People of the Mind}} 
Urban circles, chiefly in \textbf{Thepyrgos}. Scholars, civil engineers, translators of other peoples" law. \textbf{Long memories, sharp minds}; ironically, many are experts on human and Aeler culture. Their courts weigh arguments like bridges.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Everyday.} Archivists, canal designers, mint auditors, diplomatic tutors.
  ℹtem \textbf{Gifts (examples).} \emph{Memory Canticle} (recall a text with line−accurate detail); \emph{Number Music} (DV −−1 on design/repair/logistics projects framed aloud); \emph{Cold Reading} (Position +1 on first parley when you had time to observe).
  ℹtem \textbf{Strings.} gloss−tree charter; canal seal; ledger witness.
\end{itemize}

\paragraph{Sundered Elves (Itinerant Thora)} 
Lethai−thora who reject the bench for the road. They carry portable gloss−trees and teach \emph{context literacy} in market towns. Neither fully grove nor fully court, they translate between.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Gifts (examples).} \emph{Bridge−Tongue} (treat cross−cultural etiquette DV as if in shared frame once/scene); \emph{Pocket Gloss} (create a one−use Context String on the fly).
\end{itemize}

\medskipℎrule\medskip

\subsection{XV. The Two Gifts (Body & Mind) & Mixed Blood}

Lethai tradition holds: \textbf{no one may bear both gifts} at once. A person chooses (or is chosen by) \emph{Body} (Lethai−al, Lethai−ar) \emph{or} \emph{Mind} (Lethai−thora). Training and rites reinforce the choice; law and courtesy expect it.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Rule of the Two Gifts.} A Lethai PC selects either \emph{Body} or \emph{Mind} at character creation; pick Gifts only from that list. Swapping lists requires a season−long project and comes with social costs (lose one String tied to the former gift).
  ℹtem \textbf{Half−Elf Exception.} A half−elf whose heritage is \emph{quartered} (one quarter Lethai−al or −ar; one quarter Lethai−thora; remainder human or other) may take \emph{one} Body Gift and \emph{one} Mind Gift. This \textbf{Bridge−Born Clause} is rare and often controversial.
  ℹtem \textbf{Most Half−Elves.} Most half−elves in mixed cities descend from Lethai−al lines; they default to Body Gifts unless the Bridge−Born Clause applies.
\end{itemize}

\textit{In play.} The Two Gifts are \emph{narrative permissions}: Body Gifts tend to grant Position shifts and physical resistances; Mind Gifts lean toward DV reduction in planning, repair, and rhetoric. Do not add new math−−−use core ladders.

\medskipℎrule\medskip

\subsection{XVI. Hooks & Complications}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{The Gloss That Wouldn"t Read.} A treaty text from three generations back refuses to make sense. Find the missing context keys or watch a border go hot.
  ℹtem \textbf{Web−Law, River−Law.} Lethai−ar silk syndics claim a bridge−toll by Inae's charter; Lethai−thora auditors counter with canal clauses. Bowl, then Board, then \emph{Shade Etiquette}.
  ℹtem \textbf{Bridge−Born in Question.} A half−elf prodigy manifests both gifts; a circle moves to forbid the rite. Protect or persuade before the masks come down.
  ℹtem \textbf{Mask Rights Forfeit.} A serpent oath−bearer allegedly broke web−law in human courts; can a seed−tithe seal commute sanction to a repair project?
  ℹtem \textbf{The Sundered's Lesson.} A Sundered gloss−carrier teaches a market to read Lethai context−−−someone powerful profits from the confusion and wants them gone.
\end{itemize}

\medskipℎrule\medskip

\subsection{XVII. Rivals & Distance: Aeler and Ykrul}

⫕section*{With the Aeler (Stone−Law, Root−Law)}
The quarrel is old and mostly quiet: \textbf{stone−law} tallies weight and span; \textbf{root−law} tallies shade and years. Each doubts the other's ledgers. Aeler keystone stairs bite deep; Lethai−al prefer lines that bend with slope and sap. They trade proofs−−−bridges that do not drown salmon, hedges that do not choke roads−−−and keep their councils separate when tempers rise. The Aeler call elves the "Aelaef"

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Treaties in Practice.} \emph{Keystone−in−Bark} accords mark where a road enters a grove: the Aeler choose the tread, the Lethai choose the line. Breaking either creates fines paid as \emph{repair years}.
  ℹtem \textbf{Isolation Habits.} Pass−keeps that only ring bells to one side; roof−tree circles that refuse iron even when wrapped. Delegations meet at \emph{split courts}−−−half oven, half river bank−−−and adjourn before dusk.
  ℹtem \textbf{Strings.} keystone courtesy; shade−credit; river−weir writ; seed−tithe seal.
  ℹtem \textbf{Timers.} \textbf{Ridge−−Root Stalemate [4]} (project stalls until a mixed design is signed); \textbf{Sluice vs. Salmon [4]} (silt, tariffs, tempers).
  ℹtem \textbf{Levers at the Table.} If a scene honors both \emph{iron−covered} and \emph{step−on−stone}, begin in a \emph{Shared Frame} (Position +1 on the first exchange). If either party arrives with bare iron or cuts outside plan, advance \emph{Ridge−−Root Stalemate} by +1.
\end{itemize}

⫕section*{With the Ykrul (Grass−Law, Shade−Law)}
Respect edged into distance. The \textbf{Silence Furlong} keeps hooves from roots and axes from borders: cross wordless, name yourself once, and keep to the stones. When honored, councils run warm; when broken twice in a season, gray−fletched messengers appear at dusk and rings draw tight.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Customs.} \emph{Bowl, Board, and Shade}−−−fairness, then geometry, then courtesy; a concession on the Ykrul board becomes a seasonal \emph{light−due} at the forest edge.
  ℹtem \textbf{Isolation Habits.} Markets on the margin (hedge fairs) rather than deep exchange; joint patrols that HEAL hedges together but camp apart; songs traded by echo, not chorus.
  ℹtem \textbf{Strings.} light−due receipt; ferry knot; windbreak right; border−song.
  ℹtem \textbf{Timers.} \textbf{Furlong Breach [4]} (speech, grazing, or felling in the strip); \textbf{Banner at the Hedge [3]} (Meadow banners crowd the edge; tempers tighten).
  ℹtem \textbf{Levers at the Table.} When both sides keep the Furlong and the \emph{iron−covered} rule, reduce DV by −−1 on any parley for passage or pasture. On a breach, increase DV by +1 and tick \emph{Furlong Breach}.
\end{itemize}

\paragraph{Notes for Portrayal}
Keep rivalry \emph{policy−first}: quotas, stairs, hedges, ferries, and the cost of light−−−rather than eternal enmity. Isolation is a choice with reasons: fewer misunderstandings, slower trade, less shared risk. Let dissenters exist on both sides (Aeler reform lodges; Lethai merchant foresters) who argue for \emph{mixed councils} where ovens, weirs, and hedges share one ledger.

\medskipℎrule\medskip

\subsection{XVIII. War by Shade −−− Deterrence & Asymmetry}

The Lethai−al and Lethai−ar do not meet banners in open fields. They make \emph{routes expensive}, fragment columns, and end campaigns by ledger long before the first arrow. This doctrine−−−called \textbf{war by shade}−−−is why most peoples simply leave the forests alone.

\paragraph{Principles}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Fight the route, not the regiment.} Break ferries, confuse lanes, erase way−marks, and make supply walk twice.
  ℹtem \textbf{One strike, five repairs.} Every cut obligates replanting or flood−work; war ends when the ledger of repair is signed.
  ℹtem \textbf{Night, rain, understory.} Engage when wind and water cover sound; withdraw where canopy eats pursuit.
  ℹtem \textbf{Courtesy as weapon.} An iron uncovered or a crushed sprout is grounds to raise tariffs, close ferries, or call wardens without drawing a blade.
\end{itemize}

\paragraph{Plays of the Shade}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Hedge War.} Edge−wards are cut and reknit to channel intruders into \emph{dead ground} pockets watched by wardens. (Invader \emph{Move/Scout} DV +1 while \emph{Hedge Mosaic} stands.)
  ℹtem \textbf{River Denial.} Ferries vanish upstream; weirs open at dusk; mills idle to clear silt where pursuers need footing. (Start \emph{Supply Strangle [4]} on any hostile force dependent on crossings.)
  ℹtem \textbf{Night Lanes.} Lethai−ar \emph{web−law} lays silk trip−lines and warning strings; Inae's oath−bearers strike once and vanish; Isoka's keepers use venin as medicine with teeth. (First ambush in dim light begins at \emph{Dominant} for defenders; on hit, intruders mark \emph{Fatigue} instead of direct Harm unless they stand and escalate.)
  ℹtem \textbf{Canopy Runners.} Lethai−al move above sight−lines; arrows fall where footfalls never were. (Once/scene, defenders convert a \emph{pursuit} into a \emph{parley or retreat} with \emph{Position +1}.)
\end{itemize}

\paragraph{Deterrence, in Practice}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Fearsome Reputation.} Before any faction chooses a forest campaign, start \textbf{"Leave Them Alone" [2]}. On fill, leaders choose tariff, treaty, or route−around over invasion.
  ℹtem \textbf{Intruder Exhaustion [6].} Ticks for lost hours, wet powder, spoiled grain, and wrong turns. On fill, morale folds without a decisive battle.
  ℹtem \textbf{Arrow Ethics [3].} Songkeepers oversee. If intruders withdraw and sign repairs, arrows go quiet. If they burn, the mask rights come down and ambushes escalate.
\end{itemize}

\paragraph{Terrain Tags (at the table)}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Hedge Mosaic.} Invaders suffer DV +1 on Navigate/Scout; first defender action gains \emph{Position +1}.
  ℹtem \textbf{Resin Smoke.} On ignition, obscure vision and sting eyes; convert first Ranged Harm 1 against defenders to \emph{Fatigue}.
  ℹtem \textbf{Stone Steps.} If invaders \emph{step on stone} and keep iron covered, begin in a \emph{Shared Frame} (DV −−1 to request safe passage). Breaches tick \emph{Under−Root Grudge}.
\end{itemize}

\paragraph{Why Most Leave Them Alone}
\begin{itemize}[leftmargin=*]
  ℹtem Campaigns end in tariffs and repair years, not trophies−−−no glory to sell at court.
  ℹtem Columns starve by inches while ledgers fatten with fines and ferry dues.
  ℹtem Past nights have names (\emph{Two Bridges Gone, The River That Walked}); generals remember.
\end{itemize}

\paragraph{Hooks}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Five Cuts, One Bridge.} A hedge war strands a foreign regiment. Negotiate the repair ledger before someone lights the resin stores.
  ℹtem \textbf{Web & Weir.} Lethai−ar silk lines are blamed for a drowned patrol; River Denial saved three villages. Untangle oath from accident before \emph{Leave Them Alone} collapses to \emph{Banner at the Hedge}.
\end{itemize}

\medskipℎrule\medskip

\textbf{Context−Saturated Speech:} Lethai language evolves rapidly. Reading texts older than two generations requires successful Lore + Notice (DV 4−5) or a Context String. Speaking in older registers without proper context keys increases DV by +1.

\begin{center}
  {ŁARGE Lethai−ar −−− The Vowed in Silk and Scale}
  
  \bigskip
  {łarge ``Speak truth twice, whisper once; tie what you mean.''}
  \end{center}
  
  The Lethai−ar are not a separate bloodline so much as a vow. They are Lethai−al (the People of the Body, ``wood−elves'') who have sworn themselves to one of two ancient powers of the threshold: Inae, the Weaver (threads, mercy, binding context) or Isoka, the Serpent Queen (shedding, cunning, the price of cure). In some ages they are scarce, in others−−−like the present−−−their numbers grow wherever borders fray and oaths need teeth.
  
  They have shunned both the Gift of the Body and the Gift of the Mind, rejecting the paths of pure strength and pure knowledge. Instead, they seek the Blessing of Pattern from Inae and the Blessing of Change from Isoka−−−finding power in the spaces between certainty.
  
  They are the elves most other peoples prefer not to cross at night.
  
  \bigskip
  
  \textbf{Where They Live}
  
  \begin{itemize}
  ℹtem \textbf{Edge−Forests & Gloam Holds.} Rope−bridged canopy towns with silk wind−screens, scent−coded paths, and night gardens. On the valley floor: stone ``molt courts'' where the old is shed and sealed.
  ℹtem \textbf{Under−Vales & Rootways.} Knot−mapped tunnels, mirror−trick chambers, and pool−shrines where whispers travel farther than footsteps.
  ℹtem \textbf{Port Enclaves.} Lantern−draped alleys near customs houses: contract brokers, poison−cure apothecaries, and ``mask−right'' parlors where parley is held under silk.
  \end{itemize}
  
  They're rare north of the Upper Amaranthine, but cells are appearing along caravan routes and siege lines−−−wherever a binding needs mercy or a lie needs a price.
  
  \bigskip
  
  \textbf{The Vows (Two Courts, One Edge)}
  
  \textbf{The Silk Courts of Inae (Weaver)}
  \begin{itemize}
  ℹtem \textbf{Temper:} mercy with memory; the knot that mends rather than strangles.
  ℹtem \textbf{Work:} draft and enforce multi−party compacts; rebuild frayed custom; hide the vulnerable in plain sight.
  ℹtem \textbf{Signs:} three−strand cords, ledger−ribbons, masks with tear−slits.
  ℹtem \textbf{Law:} Two ledgers, Said & Meant, tied together with a visible clause.
  ℹtem \textbf{Sin:} binding without consent; mending that erases the harmed.
  \end{itemize}
  
  \textbf{The Coil Courts of Isoka (Serpent)}
  \begin{itemize}
  ℹtem \textbf{Temper:} cunning without needless cruelty; harm that carries its own cure.
  ℹtem \textbf{Work:} expose weak seams; stage molts (identity exits) for those trapped; weaponize patience.
  ℹtem \textbf{Signs:} shed−skin sashes, cup−and−vial pairs, scalpels wrapped in green thread.
  ℹtem \textbf{Law:} every poison must be paired with a cure somewhere the oath−holder can reach.
  ℹtem \textbf{Sin:} a wound with no offered remedy; a molt forced by shame.
  \end{itemize}
  
  Together: Inae binds the room to meaning; Isoka frees the person from traps. Lethai−ar make both moves feel like one motion.
  
  \bigskip
  
  \textbf{Language, Marks, and Memory}
  \begin{itemize}
  ℹtem \textbf{Context−Locked Speech.} Elven is so contextual that two generations can't read each other cleanly. Lethai−ar solve this with marks (tattoo/ink/scar/silk beadwork) that carry context−−−names of witnesses, seasons, and prices−−−on the body.
  ℹtem \textbf{Marks → Curses.} A mark without its living story curdles: a ``Hunter's Night'' becomes sleeplessness; a ``Mercy Knot'' becomes compulsion. The Lethai−ar curate and retire marks like hazardous heirlooms.
  ℹtem \textbf{Artifact Burdens.} Many sacred tools (masks, cords, vials) store cultural knowledge. Lethai−ar treat them like ledgers: if the story breaks, so does the tool's kindness.
  \end{itemize}
  
  \bigskip
  
  \textbf{The Shunned Gifts}
  
  The Lethai−ar reject two fundamental elven paths, viewing them as forms of hubris:
  
  \textbf{The Gift of the Body}  
  The pursuit of physical perfection, martial prowess, and dominance through strength. The Lethai−ar see this as a trap−−−the body ages, weakens, and fails. Those who chase only physical mastery become brittle when their bodies betray them.
  
  \textbf{The Gift of the Mind}  
  The worship of pure knowledge, magical power, and intellectual supremacy. The Lethai−ar view this as dangerous isolation−−−the mind can become lost in abstractions, disconnected from the web of relationships that bind communities together.
  
  Instead, they seek the \textbf{Blessing of Pattern} from Inae and the \textbf{Blessing of Change} from Isoka−−−gifts that come not from individual perfection, but from understanding one's place in the greater weave of existence.
  
  \bigskip
  
  \textbf{Etiquette & Rites (How to Survive Their Parlors)}
  \begin{itemize}
  ℹtem \textbf{Mask−Right.} Serious parley is held under silk; roles and intentions are declared, then the mask goes on. Break mask−right and you will not be heard again that season.
  ℹtem \textbf{Speak Twice, Whisper Once.} State the truth two ways; then the price as a whisper to the witness. If you can't name the price, you don't understand your own request.
  ℹtem \textbf{Thread Before Blade.} Offer a binding solution first. If steel must speak, it speaks where a thread was refused.
  ℹtem \textbf{Vial Courtesy.} In Isoka's houses, a dose sits beside its antidote. Taking one without the other marks you as untrustworthy.
  \end{itemize}
  
  \bigskip
  
  \textbf{War Without Battle (Doctrine)}
  
  The Lethai−ar are the textbook on asymmetric warfare:
  \begin{itemize}
  ℹtem \textbf{Silence Corridors.} Silence furlongs across borders: no grazing, no felling, no names. Violations turn your scout signs against you.
  ℹtem \textbf{Ambush as Audit.} They treat logistics like law−−−cut one keystone (a ferry, a ridge path) and make you pay repair years under witness.
  ℹtem \textbf{Fear Economy.} Their reputation buys exits before arrows. Most armies learn to avoid ``black ribbon nights,'' when coils and webs both shift.
  ℹtem \textbf{Night Doctrine.} Lamp placements create ``honor lanes'' where even pursuers must slow. They kill lights, not people−−−unless you force the price.
  \end{itemize}
  
  Most peoples leave them alone; those who don't tend to meet a contract they wish had been a skirmish.
  
  \bigskip
  
  \textbf{Neighbors & Borders}
  \begin{itemize}
  ℹtem \textbf{Ykrul:} Mutual distance, ritual respect. Border treaties often include a speechless strip and a joint ``Bowl & Board'' at season's end. Ykrul price routes; Lethai−ar price behavior.
  ℹtem \textbf{Aeler:} Underway adjacency breeds friction. Aeler bring breath−time and beams; Lethai−ar bring mask−right and mercy−knots. Both enforce oaths; they disagree about who writes the receipt.
  ℹtem \textbf{Lethai−al / Lethai−thora:} The Lethai−ar recruit from the al; they contest with the thora over who defines ``context.'' Sundered thora wanderers sometimes find refuge under coil or web.
  \end{itemize}
  
  \bigskip
  
  \textbf{Economy & Everyday}
  \begin{itemize}
  ℹtem \textbf{Web−Law Brokerage.} Multi−party deals that hold even across language drift.
  ℹtem \textbf{Poison−Cure Houses.} Venoms for cutting cruelty short; antidotes for those who choose return.
  ℹtem \textbf{Night Markets.} Silk−screened stalls that sell contexts: a role, a mask, a witness−for−hire who'll remember what you meant.
  \end{itemize}
  
  \bigskip
  
  \textbf{What They Believe They're For}
  \begin{itemize}
  ℹtem To bind harm so it cannot spread.
  ℹtem To cut traps that call themselves tradition.
  ℹtem To carry context so tomorrow's children won't inherit our mistakes as their curses.
  \end{itemize}
  
  They will tell you, if you ask in the right room: ``We are the price you pay to keep meaning.''
  
  \bigskip
  
  \textbf{Frictions & Story Heat}
  \begin{itemize}
  ℹtem \textbf{Marks Going Bad.} A famous mark has lost its last living witness; now it behaves like a curse. Do you retire it−−−or repair the story it needs?
  ℹtem \textbf{Mask−Right vs. Under−Beam.} Aeler want a dispute under a single lamp (witness law). Lethai−ar demand silk (role law). Which threshold rules?
  ℹtem \textbf{Snake's Mercy.} A tyrant's guard captain begs a molt (new life) after unspeakable work. The cure exists−−−who pays for it?
  ℹtem \textbf{Thread in the Steppe.} A silence furlong through Ykrul pasture is broken; wolves and politics arrive together. Re−stitch the corridor or accept a border that now bites.
  \end{itemize}
  
  \bigskip
  
  \textbf{How to Portray Them (at the table)}
  \begin{itemize}
  ℹtem \textbf{Embodied calm.} Even at war, they stand like dancers waiting for the beat.
  ℹtem \textbf{Context first.} Every request has a price, every truth needs a witness, every blade wants a thread it can't cut.
  ℹtem \textbf{Mercy with edge.} They save people by binding situations−−−and they do it with the precision of surgeons, not saints.
  \end{itemize}
  
  The Lethai−ar do not hide because they are weak. They hide because the world forgets context quickly, and they would rather HEAL it than burn it. By shunning the Gifts of Body and Mind, they seek a deeper wisdom−−−the Blessing of their chosen patrons, earned through submission to pattern and embrace of necessary change.
  
  The Lethai−ar are the elves you negotiate with if you're wise−−−and fight only if you enjoy losing to rules you didn't bother to learn.
    
\subsection{XIX. Marks −−− Burden, Blessing, and Lost Context}

Many Lethai artifacts carry \textbf{inherited knowing}. Leafblades etched with moss−tallies, resin−glass panels that hold songs, riverstones scored by ferrymen−−−these do not merely \emph{remember}; they \emph{obligate}. Among the Lethai−−−especially the woodwise Lethai−al−−−\textbf{past knowledge is a burden on the present}. What you inherit, you must carry carefully or set down with witness.

\paragraph{Passing−Knowledge Artifacts}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Leafblade (moss−tallied).} A knife whose edge carries hedge−law. \emph{Invoke:} gain a \emph{Context String} for border disputes; \emph{Cost:} tick \textbf{Burden Ledger [1]} (you owe repair work this season).
  ℹtem \textbf{Resin−Glass Pane.} Sun−cured panels that hold a chorus. \emph{Invoke:} reduce DV by −−1 on rites requiring the old wording; \emph{Cost:} \emph{Under−Root Grudge} +1 if used to excuse harm without replanting.
  ℹtem \textbf{Riverstone Book.} Pebbles in corded order record ferry rights. \emph{Invoke:} create a temporary \emph{Ferry Knot String}; \emph{Cost:} pay \emph{light−due} before you drink.
\end{itemize}

\paragraph{Marks of the Body (Animist Rites)}
Lethai \textbf{Marks} (ink, scar, resin−inlay) amplify embodied gifts by binding person to pattern. Every Mark has \emph{Context Keys}−−−place, season, witness. \textbf{Out of context, a Mark drifts toward curse.}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Barkskin Mark.} Skin takes the grain of oak. \emph{Gift:} once/scene, downgrade a Physical Harm by one step (\emph{Position +1} when braced). \emph{Context Keys:} shaded work, sap rite, roof−tree witness. \emph{Curse (lost context):} brittleness−−−cold converts first Fatigue to Harm.
  ℹtem \textbf{River−Ears.} Cartilage braided with silver reed. \emph{Gift:} sense current/voices across water (Clue +1 on rivers). \emph{Keys:} pour water first, ferryman's knot. \emph{Curse:} flood−whispers−−−compelled toward lowest ground (SB: \emph{Downhill Pull}).
  ℹtem \textbf{Spider's Patience} (Inae oath). \emph{Gift:} hold still beyond breath; ambush begins at \emph{Dominant} in dim light. \emph{Keys:} web−tithe, oath−sign shown. \emph{Curse:} fixation−−−DV +1 to abandon a plan once set.
  ℹtem \textbf{Serpent's Breath} (Isoka oath). \emph{Gift:} once/scene convert Toxin Harm 1 to Fatigue; read heat like color. \emph{Keys:} antivenin rite, shed−skin offering. \emph{Curse:} cold−blooded−−−action slows in chill (Position −−1 unless warmed).
  ℹtem \textbf{Storm−Shadow.} Ink of ash and rain. \emph{Gift:} vanish in downpour; ranged attacks vs. you suffer −−1 Effect in rain. \emph{Keys:} storm cup, gutter song. \emph{Curse:} thunder−call−−−loud noises pull attention (start \emph{Noticed [1]} on loud scenes).
  ℹtem \textbf{Lanternbone.} Resin set in a healed break. \emph{Gift:} glow faintly to mark safe steps; \emph{once/scene} grant ally \emph{Position +1} on footing checks. \emph{Keys:} night watch, seed−tithe. \emph{Curse:} beacon−−−predators test the light (GM may spend SB to start \emph{Hungry Eyes [2]}).
\end{itemize}

\paragraph{Curses as Lost Context}
A Mark becomes a \textbf{Curse} when used without its Keys or against its ethic.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Context Drift [4].} Ticks when a Mark is invoked off−season, off−place, or without witness. On fill, the Mark expresses its \emph{Curse} until \emph{Context is Restored}.
  ℹtem \textbf{Context Restored (project [4]).} Recover Keys: return to place, invite songkeeper, repay light−due. Aelinnel \emph{Sumwright gloss−trees} can substitute for witness once.
  ℹtem \textbf{Burden Ledger [6].} Tracks obligation from inherited knowledge. On fill, pay with repair years, seed−tithes, or forfeit a String tied to the artifact.
\end{itemize}

\paragraph{At the Table (simple levers)}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Strings:} \emph{mark−right}, \emph{context key}, \emph{songkeeper witness}, \emph{gloss−tree charter}. Spend to reduce DV by −−1 on rites/negotiations tied to that Mark or artifact.
  ℹtem \textbf{Position/Effect:} Most Marks grant \emph{Position +1} or \emph{Effect +1} \emph{once/scene} when used with Keys; without Keys, instead tick \emph{Context Drift}.
  ℹtem \textbf{Mitigation:} Aeler \emph{null−bells} can mute a Mark's \emph{Curse} for a scene at the cost of \emph{lamp−time} or \emph{Public Outrage [1]} if used in a Lethai court.
\end{itemize}

\paragraph{Custodians & Dissent}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Songkeepers} steward Keys and judge when burdens may be set down.
  ℹtem \textbf{Lethai−ar Oathbinders} police oath−Marks; mask rights fall for abusers.
  ℹtem \textbf{Lethai−thora Archivists} map Marks to contexts in \emph{gloss−trees}; Sundered Elves carry these into markets and mistakes.
\end{itemize}

\paragraph{Hooks}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{The Bark That Broke.} A famed Barkskin bearer shattered in frost−−−find the missing roof−tree witness before winter claims more.
  ℹtem \textbf{River−Ears in a Dry City.} A collector misuses a Riverstone Book; canals riot. Restore Keys or rewrite dues with a bowl at the fountain.
  ℹtem \textbf{Spider's Ledger.} Inae's Mark is called curse after a web−judge froze a harvest lane. Untangle oath from context drift before masks come down.
\end{itemize}

\section{Valewood −−− Empire Under Leaves}

\subsection*{Living courts, phasing ruins, and the law of guest and host}

\begin{quote}
"Name yourself once, pour water first, and step on the stones. In the Valewood, truth grows wild and the paths remember your feet."
\end{quote}

\medskipℎrule\medskip

\subsection{I. What the Valewood Is}
A vast, old−growth forest whose memory runs deeper than city annals. Phasing ruins (\emph{star−roads, living stone, breathing streets}) surface and sink with moon and season; oaths are measured in years of repair. Power moves through \textbf{hospitality}, \textbf{knowledge of ways}, and \textbf{careful speech}. The Lethai−al keep edge−wards and ferry rights; the Lethai−ar hold \emph{web−law} and shaded corridors; fae courts barter in \emph{truth told right now}; beast−kin patrol cairns none can map twice the same way.

\medskipℎrule\medskip

\subsection{II. Courtesy & Entry (Shade Etiquette)}
Observe these to avoid waking the wood's ledgers.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Iron Covered.} Wrap iron in leather or cloth. \emph{(Position +1 on first parley at a hedge or ferry.)}
  ℹtem \textbf{Name Once.} Speak your name and intent at the edge; do not name the forest as property.
  ℹtem \textbf{Step on Stone.} Use set stones and boardwalks; crushed sprouts become \emph{Light−Due}.
  ℹtem \textbf{Water First.} Pour the first cup to the cistern or stream; then drink.
  ℹtem \textbf{Leave the Light.} Replace what you take: plant, HEAL, or pay \emph{seed−tithe}.
\end{itemize}
\textit{Timers for breaches:} \textbf{Under−Root Grudge [4]} (tariffs, closed ferries), \textbf{Stream−Clouding [4]} (mill shutdowns, fish wardens).

\medskipℎrule\medskip

\subsection{III. Routes & Taboos (Using the Travel Seed)}
\paragraph{Shadow Corridor (Thin Shore).} The misted coastal lane north−−south toward Zakov. Draw \emph{Spade} from Valewood, \emph{Heart} from Mistlands (or Valewood), \emph{Club} from Mistlands (wraiths, bell−line failures), \emph{Diamond} from either Mistlands (\emph{Ward−salt, Lantern Writ}) or Valewood (\emph{Way−Cord, Truce−Bough}) depending on which law you invoke.

\paragraph{Rule of 9s (Valewood & Theona).} Whenever a 9 appears in a travel seed touching Valewood or the isle moots, add an \textbf{omission}: a missing step, an unsaid name, an unseen guest. If the 9 is a \emph{Diamond}, you may break the taboo once−−−someone will come to collect. Treat the omission as a one−scene \emph{Shared Complication} the table must name.

\medskipℎrule\medskip

\subsection{IV. Places That Hold (Spade Prompts)}
Use as scenes, sets, or advantages:
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Star−Road Shard.} Pale flagstones hum when trod in order; a gate stirs if you match the cadence.
  ℹtem \textbf{Rooted Amphitheatre.} Moss seats remember speeches not yet given; debate here shapes routes.
  ℹtem \textbf{Glyphed Bridge.} Lies make glyphs glow; truth pays the toll.
  ℹtem \textbf{Calendar Grove.} Solstice−aligned trunks; one tree is stone−−−time tallies are read here.
  ℹtem \textbf{Unfound Arcade.} Arches appear at dusk, vanish at dawn; reality keeps hours.
  ℹtem \textbf{Breathing City.} Vine−choked streets shift on the hour; the ruin is alive and listening.
\end{itemize}

\medskipℎrule\medskip

\subsection{V. Pressures of the Wood (Club Prompts)}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Ward−Trap.} Ancient mechanisms treat you like yesterday's invader; tools and talk jam.
  ℹtem \textbf{Oath−Magnet.} Words stick; a careless promise rewrites your plan (\emph{start a Task Timer you named aloud}).
  ℹtem \textbf{Geas Catch.} A phrase turns into a job that names itself; the wood expects it done.
  ℹtem \textbf{Name−Theft.} Edges of your true name are sampled; masks come down unless paid in courtesy.
  ℹtem \textbf{Muster of Boughs.} Green banners rise−−−travel becomes trespass until a \emph{Truce−Bough} is raised.
  ℹtem \textbf{Mirror Rain.} Leaf−drips replay choices you didn"t make; doubt taxes Position until resolved.
\end{itemize}

\medskipℎrule\medskip

\subsection{VI. Charms, Rights, & Papers (Diamond Prompts)}
Treat these as \emph{Strings} that spend like documents or single−use boons:
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Way−Cord.} Knot that points to the true next turn once. \emph{Spend:} DV −−1 on Navigate/Scout; negates one \emph{Sweet Wind} lie.
  ℹtem \textbf{Dew−Mirror.} Shows a thing as it is, not as sung, for one scene. \emph{Spend:} reveal a concealed intent or hazard; ignore one glamour.
  ℹtem \textbf{Hazel Token.} Lawful crossing of a warded hedge without snag. \emph{Spend:} Position +1 to pass an edge−ward.
  ℹtem \textbf{Honey−Right.} Eat and speak under bee−stone protection. \emph{Spend:} immunity to first social SB this parley.
  ℹtem \textbf{Name−Bead.} Kept promise warms, broken one chills and glows. \emph{Spend:} convert one consequence into \emph{Repair Work [4]}.
  ℹtem \textbf{Truce−Bough.} Cut, hunt, or camp in a named copse without offense. \emph{Spend:} suppress \emph{Muster of Boughs} for a scene.
  ℹtem \textbf{City−Shard.} Wake one gate or stair in a phasing ruin. \emph{Spend:} create an exit where none is seen.
  ℹtem \textbf{Oathsap.} Seal a pact that even echoes respect. \emph{Spend:} turn a verbal concession into a season−long right.
\end{itemize}

\medskipℎrule\medskip

\subsection{VII. People & Practice}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Pathweavers.} Guides who knot safe routes into cord; their cords double as minutes of what was promised on the way.
  ℹtem \textbf{Wardens of the Roof−Tree.} Canopy surveyors and hedge−keepers; they measure damage in \emph{repair years}.
  ℹtem \textbf{Silk Engineers (Lethai−ar).} Keep night lanes and warning webs; \emph{mask rights} fall for oath abuse.
  ℹtem \textbf{Songkeepers.} Render gloss and context for old speech; without them, even yesterday's syntax can betray.
  ℹtem \textbf{Beast−Kin.} Border guardians tied to cairns and seasons; they bargain in food, paths, and respect.
\end{itemize}

\medskipℎrule\medskip

\subsection{VIII. War by Shade (Deterrence in the Valewood)}
The wood teaches \emph{route warfare}: make supply walk twice; end campaigns by ledger. Ambushes begin at \emph{Dominant} under rain or understory; invaders who \emph{step on stone} and keep iron covered may parley for escorted withdrawal instead of bleed for it. Before anyone plans a forest campaign, start \textbf{"Leave Them Alone" [2]}; on fill, command chooses tariff, treaty, or route−around.

\medskipℎrule\medskip

\subsection{IX. Integration Hooks}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Political Intrigue.} \emph{Bowl, Board, and Shade}: fairness, then geometry, then courtesy. Convert a concession into a \emph{season−long right} at a ford, gate, or pasture.
  ℹtem \textbf{Caravans.} Shadow Corridor runs require \emph{Way−Cord} or \emph{Lantern Writ}. Mistland law (ward−salt) vs. Valewood law (truce−bough) determines your final judgment.
  ℹtem \textbf{Wilderness.} Treat edge−wards and fuel mosaics as special orders during camp phases; blight quarantine uses \emph{ash−wash stations}.
  ℹtem \textbf{Aeler Roads.} \emph{Keystone−in−Bark} accords: Aeler choose tread, Lethai choose line. Breaking either creates fines paid as \emph{repair years}.
  ℹtem \textbf{Ykrul Borders.} \emph{Silence Furlong}: cross wordless, then speak once. Keep it and reduce DV −−1 on pasture passage; break it and tick \emph{Furlong Breach}.
\end{itemize}

\medskipℎrule\medskip

\subsection{X. Omens & Seeds}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Sweet Wind.} A breeze that lies kindly; \emph{Wind−Veil} mutes it for a walk.
  ℹtem \textbf{Missing Ninth.} Somewhere a ninth step or name is absent; pay in favors or get lost kindly.
  ℹtem \textbf{Bell−Line Failure.} Mistland lantern law flickers; wraith indemnities come due.
\end{itemize}

\paragraph{Adventure Seeds}
\begin{enumerate}[leftmargin=*]
  ℹtem \textbf{The Bowl at the Bridge.} A \emph{Glyphed Bridge} lights at every lie; arbitrate a toll war using \emph{Bowl, Board, and Shade}.
  ℹtem \textbf{Way−Cord for a Price.} A pathweaver demands a \emph{Name−Bead} to knot your route; who pays when it chills?
  ℹtem \textbf{Empire Wakes.} A \emph{Breathing City} aligns at dusk; open a \emph{City−Shard} gate or be counted as yesterday's invader.
  ℹtem \textbf{Silk and Sluice.} Lethai−ar web−law snared a patrol; river pilots claim the weir saved three villages. Balance oath and repair.
\end{enumerate}

\medskipℎrule\medskip

\subsection{XI. At the Table (Simple Levers)}
Keep to core ladders−−−use these as narrative permissions:
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Context Strings} (\emph{gloss−tree, songkeeper witness}) reduce DV −−1 when reading/pleading in older registers.
  ℹtem \textbf{Light−Due & Seed−Tithe} convert fines into \emph{Repair Projects [4]} instead of exile or coin.
  ℹtem \textbf{Way−Cord / Truce−Bough / Dew−Mirror} spend for single−scene Position/DV shifts as listed; abusing them starts \emph{Under−Root Grudge}.
\end{itemize}

\textit{Portrayal note:} Let policy speak before mystique−−−ferries, hedges, quotas, and the cost of light−−−then add the wonder: paths that hum, rain that remembers, courts that sit where the river can overhear.

\section*{Starting Talents}

\subsection*{Lethai−al/ar Path: Embodied Presence}
\talent{Heightened Senses (3 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Wits 2+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to Notice and Survival rolls
      ℹtem Can detect hidden creatures/objects with successful Wits + Notice (DV 3)
      ℹtem In natural environments, +1 Position on stealth and tracking rolls
    \end{itemize}
\end{itemize}

\talent{Root−Balance (3 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Body 2+, Heightened Senses
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to Athletics and resist shove/knockdown attempts
      ℹtem Can move through natural terrain without penalty
      ℹtem Once per scene, can stabilize on precarious footing with a successful Body + Athletics roll (DV 3)
    \end{itemize}
\end{itemize}

\subsection*{Lethai−thora Path: Mental Acuity}
\talent{Long Memory (3 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Spirit 2+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem Perfect recall of events within the past week
      ℹtem +1 die to Lore and Insight rolls involving historical or cultural knowledge
      ℹtem Once per session, can recall a crucial detail from long−term memory (GM's discretion)
    \end{itemize}
\end{itemize}

\talent{Cold Reading (3 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Wits 2+, Long Memory
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to Sway and Insight rolls involving observation and social analysis
      ℹtem Can make intuitive leaps about people's motivations with successful Wits + Insight (DV 3)
      ℹtem Once per scene, gain +1 Position on first parley when you had time to observe
    \end{itemize}
\end{itemize}

\section*{Key Talents}

\subsection*{Lethai−al Specific}
\talent{Canopy Spring (4 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Heightened Senses, Body 2+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to climbing, leaping, and arboreal movement
      ℹtem Can perform feats of agility that would normally be impossible with a successful Body + Athletics roll (DV 4)
      ℹtem Once per scene, gain +1 Position on movement actions in forested environments
    \end{itemize}
\end{itemize}

\talent{Scent of Rain (4 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Heightened Senses, Survival 1+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to weather prediction and detecting environmental hazards
      ℹtem Can sense approaching storms, fires, or diseases with successful Wits + Survival (DV 3)
      ℹtem In natural environments, can track by scent trails with +1 Effect
    \end{itemize}
\end{itemize}

\subsection*{Lethai−thora Specific}
\talent{Memory Canticle (4 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Long Memory, Lore 2+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem Can recall texts with line−accurate detail
      ℹtem +1 die to research and translation rolls
      ℹtem Once per scene, can provide crucial historical context that advances investigation
    \end{itemize}
\end{itemize}

\talent{Number Music (4 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Long Memory, Wits 2+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to design, repair, and logistics projects
      ℹtem Can perform complex calculations and engineering analysis with Wits + Craft (DV 3)
      ℹtem When framing projects aloud, reduce DV by −1 on related rolls
    \end{itemize}
\end{itemize}

\subsection*{Lethai−ar Specific (Oathbound)}
\talent{Weaver's Reading (4 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Heightened Senses, Bond to Inae
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to uncovering plots, paths, and hidden networks
      ℹtem Can sense web−based communication and surveillance systems
      ℹtem Once per scene, trace a network connection with successful Wits + Notice (DV 3)
    \end{itemize}
\end{itemize}

\talent{Venin Lore (4 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Bond to Isoka, Survival 1+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to identifying and treating poisons
      ℹtem Can convert Harm 1 from toxins to Fatigue once per scene
      ℹtem Can craft antidotes and venins with successful Lore + Survival (DV 4)
    \end{itemize}
\end{itemize}

\talent{Stillness (5 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Weaver's Reading OR Venin Lore, Stealth 2+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to stealth and hiding rolls in dim light
      ℹtem Can remain unnoticed until you move to act
      ℹtem Once per scene, become effectively invisible in appropriate cover for one exchange
    \end{itemize}
\end{itemize}

\subsection*{Sundered Elves Specific}
\talent{Bridge−Tongue (4 XP − Minor Talent):}
\begin{itemize}
  ℹtem Requirements: Long Memory, Sway 2+
  ℹtem Benefits:
    \begin{itemize}
      ℹtem +1 die to cross−cultural communication and translation
      ℹtem Can treat cross−cultural etiquette DV as if in shared frame once per scene
      ℹtem Once per session, create a one−use Context String on the fly
    \end{itemize}
\end{itemize}

\subsection*{Half−Elf Exception}
\talent{Bridge−Born Clause (6 XP − Major Talent):}
\begin{itemize}
  ℹtem Requirements: Half−elf heritage, both Lethai−al and Lethai−thora lineage
  ℹtem Benefits:
    \begin{itemize}
      ℹtem Can access both Gift of the Body and Gift of the Mind
      ℹtem +1 die to rolls combining physical and mental skills
      ℹtem Once per scene, can perform actions requiring both gifts simultaneously
      ℹtem Immune to The Curse of Division
    \end{itemize}
\end{itemize}

\section*{Cultural Mechanics}

\subsection*{Shade Etiquette}
Lethai−al courtesy that grants hospitality and hearing:
\begin{itemize}
  ℹtem \textbf{Iron Covered:} Bare iron offends ward−lines; wrap it in leather or cloth. (On first entry, covered iron grants Position +1 in parley)
  ℹtem \textbf{Name Once:} Speak your name and intent at the edge; do not name the forest as property
  ℹtem \textbf{Step on Stone:} Where stones are laid, use them; every crushed sprout is a debt
  ℹtem \textbf{Water First:} Pour a first cup to the river or cistern before you drink
  ℹtem \textbf{Leave the Light:} Replace the shade you take: plant, HEAL, or pay
\end{itemize}

\subsection*{Light−Due System}
Light and shade are currencies. Every fell, ferry, and fire has a light−due−−−balance to be repaid in replanting, canal−clearing, or trade at kinder rates.
\begin{itemize}
  ℹtem Cash a light−due to gain DV −1 on operations framed as repair, replanting, or flood work
  ℹtem Abuse it and start Under−Root Grudge [1]
\end{itemize}

\subsection*{Context Keys}
Lethai speech requires context keys:
\begin{itemize}
  ℹtem Place−name, season−mark, kinship−hand, and roof−tree sign
  ℹtem Absent any two, assume misreadings
  ℹtem When reading or pleading in older registers, increase DV by +1 unless holding a Context String
  ℹtem Cashing a Context String reduces DV by −1 and creates a Shared Frame (Position +1 on first exchange)
\end{itemize}

\section*{Strings (Lethai Cultural Influence)}
\begin{itemize}
  ℹtem \stringtype{Light−Due Receipt:} Balance to be repaid in environmental work
  ℹtem \stringtype{Shade−Credit:} Favor with forest communities
  ℹtem \stringtype{Ferry Right:} Priority passage on waterways
  ℹtem \stringtype{Resin Share:} Access to valuable forest products
  ℹtem \stringtype{Canoe−Lane Priority:} Preferred river routes
  ℹtem \stringtype{Seed Tithe:} Right to harvest forest reproduction
  ℹtem \stringtype{Spider−Vow Token:} Oathbound commitment marker
  ℹtem \stringtype{Serpent Oath:} Isoka−bound promise
  ℹtem \stringtype{Silk−Tithe:} Payment for web−based services
  ℹtem \stringtype{Gloss−Tree Charter:} Authority to interpret ancient texts
  ℹtem \stringtype{Canal Seal:} Waterway management rights
  ℹtem \stringtype{Ledger Witness:} Authority in scholarly disputes
  ℹtem \stringtype{Keystone Courtesy:} Aeler−Lethai cooperation agreements
  ℹtem \stringtype{River−Weir Writ:} Dam and water control permissions
  ℹtem \stringtype{Windbreak Right:} Ykrul−Lethai border accommodations
  ℹtem \stringtype{Border−Song:} Cross−cultural musical agreements
\end{itemize}

\section*{Patron Relationships}
\begin{itemize}
  ℹtem \textbf{Lethai−al:} Often bond with nature−related Patrons (Old Man of the Black Forest, Carrion King, Nidhoggr)
  ℹtem \textbf{Lethai−thora:} Prefer knowledge−related Patrons (The Witness, Sacred Geometry, Clockwork Monad)
  ℹtem \textbf{Lethai−ar:} Bond with Inae (spiders) or Isoka (serpents) − dark, transformative Patrons
  ℹtem \textbf{Half−elves:} Can bond with any Patron but often struggle with divided loyalties
\end{itemize}

\section*{Complications}
⨾licationtype{The Curse of Division:}
\begin{itemize}
  ℹtem When attempting to use talents from both paths, suffer −1 die to all rolls until scene ends
  ℹtem Can only advance in one path per tier without the Bridge of Gifts talent
\end{itemize}

⨾licationtype{Contextual Communication:}
\begin{itemize}
  ℹtem Cannot effectively communicate with other Lethai without proper context keys
  ℹtem Modern elven speech confuses outsiders (−1 die to social rolls with non−elves)
  ℹtem Reading old texts requires special effort or assistance
\end{itemize}

⨾licationtype{Cultural Tension:}
\begin{itemize}
  ℹtem Lethai−al and Lethai−thora communities often distrust each other
  ℹtem Half−elves face prejudice from both communities
  ℹtem Lethai−ar are feared and often ostracized
\end{itemize}

⨾licationtype{Environmental Responsibility:}
\begin{itemize}
  ℹtem Every action in natural environments has consequences
  ℹtem Breaking Shade Etiquette triggers social and environmental backlash
  ℹtem Light−dues accumulate and must be repaid
\end{itemize}

\section*{Sample Characters}

\subsection*{Talan of Reedfall, Who Unknotted the Flood}
\begin{itemize}
  ℹtem Body 3, Wits 3, Spirit 2, Presence 2
  ℹtem Skills: Athletics 2, Survival 2, Notice 1, Craft 1
  ℹtem Talents: Heightened Senses, Root−Balance, Canopy Spring, Scent of Rain
  ℹtem Affinity: Gift of the Body
  ℹtem Strings: Ferry Right, Resin Share, Shade−Credit
  ℹtem Complication: Environmental Responsibility
\end{itemize}

\subsection*{Mira of the High Archive}
\begin{itemize}
  ℹtem Body 2, Wits 4, Spirit 3, Presence 2
  ℹtem Skills: Lore 3, Insight 2, Sway 1, Craft 1
  ℹtem Talents: Long Memory, Cold Reading, Memory Canticle, Number Music
  ℹtem Affinity: Gift of the Mind
  ℹtem Strings: Gloss−Tree Charter, Ledger Witness, Canal Seal
  ℹtem Complication: Cultural Tension
\end{itemize}

\subsection*{Sariel Half−Elf of the Crossroads}
\begin{itemize}
  ℹtem Body 3, Wits 3, Spirit 2, Presence 3
  ℹtem Skills: Melee 2, Lore 2, Survival 1, Sway 1
  ℹtem Talents: Heightened Senses, Long Memory, Bridge−Born Clause, Bridge−Tongue
  ℹtem Affinity: Both Gifts (through Bridge−Born Clause)
  ℹtem Strings: Shade−Credit, Ferry Right, Gloss−Tree Charter
  ℹtem Complication: Cultural Tension
\end{itemize}

\section*{Adventure Hooks}
\begin{enumerate}
  ℹtem \textbf{The Hedge That Ate a Road:} An edge−ward grew too well and swallowed a toll lane. Broker a cut and a replanting that keeps both law and trade.
  ℹtem \textbf{The Gloss That Wouldn't Read:} A treaty text from three generations back refuses to make sense. Find the missing context keys or watch a border go hot.
  ℹtem \textbf{Web−Law, River−Law:} Lethai−ar silk syndics claim a bridge−toll by Inae's charter; Lethai−thora auditors counter with canal clauses. Bowl, then Board, then Shade Etiquette.
  ℹtem \textbf{Bridge−Born in Question:} A half−elf prodigy manifests both gifts; a circle moves to forbid the rite. Protect or persuade before the masks come down.
  ℹtem \textbf{Blight Under Boots:} Bark Blight rides a caravan's wheel hubs. Build ash−wash stations without sparking a tariff war.
\end{enumerate}

\section*{Integration with Core Rules}

\subsection*{The Curse Mechanic}
The Curse of Division works as a persistent condition:
\begin{itemize}
  ℹtem When a Lethai attempts to use talents from both paths in the same scene, mark 1 segment on a Curse Timer [4]
  ℹtem When the timer fills, suffer −1 die to all rolls until the next dawn
  ℹtem The Bridge−Born Clause talent negates this effect
\end{itemize}

\subsection*{War by Shade}
Lethai defensive warfare principles:
\begin{itemize}
  ℹtem \textbf{Hedge War:} Edge−wards channel intruders into dead ground pockets (Invader Move/Scout DV +1)
  ℹtem \textbf{River Denial:} Ferries vanish, weirs open, mills idle (Start Supply Strangle [4] on hostile forces)
  ℹtem \textbf{Night Lanes:} Web−law trip lines, single strikes, venom as medicine (First ambush in dim light begins at Dominant)
  ℹtem \textbf{Canopy Runners:} Movement above sight−lines, arrows from impossible angles (Once/scene, convert pursuit to parley/retreat with Position +1)
\end{itemize}

\subsection*{Terrain Tags}
\begin{itemize}
  ℹtem \textbf{Hedge Mosaic:} Invaders suffer DV +1 on Navigate/Scout; first defender action gains Position +1
  ℹtem \textbf{Resin Smoke:} On ignition, obscure vision and sting eyes; convert first Ranged Harm 1 against defenders to Fatigue
  ℹtem \textbf{Stone Steps:} If invaders step on stone and keep iron covered, begin in Shared Frame (DV −1 to request safe passage)
\end{itemize}
\clearpage

\begin{center}
  {ŁARGE Ge'hai −−− Those Who Carry the Edge}
  
  \bigskip
  {łarge Lethai elite cadres in body and in mind}
  \end{center}
  
  Among the Lethai, ge'hai is not a rank so much as a charge: to carry the edge where the people meet the world. The title is awarded in moot by three witnesses−−−one of body, one of mind, and one of place (a grove, a hall, a lamp under beam). From that day a ge'hai belongs to the work more than to themselves.
  
  Two great traditions shape them:
  \begin{itemize}
  ℹtem \textbf{Lethai−al ge'hai (``People of the Body'')} −−− canopy commandos and keystone saboteurs, feared for asymmetric warfare and the calm speed with which they unmake an enemy's plan.
  ℹtem \textbf{Lethai−thora ge'hai (``People of the Mind'')} −−− court tacticians and context−binders, infamous for turning rooms with etiquette, syntax, and the careful placement of witnesses.
  \end{itemize}
  
  Both hold the same maxim: \textbf{thread before blade}. Bind the harm; if binding fails, cut cleanly.
  
  \bigskip
  
  \section*{I. Lethai−al Ge'hai −−− Bark & Sinew}
  
  \textbf{Doctrine:} Silence, keystone, vanish. Lethai−al ge'hai do not ``win battles''; they remove the pieces battles would need−−−ferries, fodder, bridges, nerve.
  
  \textbf{Training:} Night−sight drills, breath discipline, canopy movement, keystone reading (how a structure stands), mark−keeping (context tattoos that store orders and prices), and silence furlong protocol (no names, no smoke, no steel until witness).
  
  \textbf{Typical cadre (6−−12):}
  \begin{itemize}
  ℹtem \textbf{Path−Cutter} (route denial; hedge & rope),
  ℹtem \textbf{Keystone−Reader} (ferries, arches, dam lips),
  ℹtem \textbf{Lantern−Thief} (light management; moon−mirrors),
  ℹtem \textbf{Molt−Keeper} (field medic; poison/antidote pairing),
  ℹtem \textbf{Whisper−Sergeant} (hand−sign command),
  ℹtem \textbf{Sable−Bow} (harassment, lure, pursuit break).
  \end{itemize}
  
  \textbf{Field rites & etiquette}
  \begin{itemize}
  ℹtem \textbf{Mask−Right (field):} canvas veils with inked roles; parley can be held anywhere the masks are donned.
  ℹtem \textbf{Vial Courtesy:} dose and cure travel together; to harm without an offered remedy is a ge'hai sin.
  ℹtem \textbf{Speak Twice:} the order is given in common speech and then once in sign; if they diverge, the sign holds.
  \end{itemize}
  
  \textbf{Operations menu (what they actually do):}
  \begin{itemize}
  ℹtem \textbf{Keystone Audit:} Mark the one stone, beam, or rope that bears a route. Remove or HEAL as needed; leave a repair year ledger if you break it.
  ℹtem \textbf{Lamp Game:} Extinguish all but one light to create an honor lane. Pursuers must slow or lose face.
  ℹtem \textbf{Gardened Ambush:} Shape a killing ground with vines, wire, and scent; fire once, then disappear into the furlong.
  ℹtem \textbf{Ferry Night:} Hold a river crossing with bell−tokens and chalked prices; turn a rout into an orderly retreat.
  \end{itemize}
  
  \textbf{Strings / Diamonds they carry:}
  \begin{itemize}
  ℹtem \textbf{Context Key} (treat a warded path as neutral once),
  ℹtem \textbf{Mask−Right Seal} (declare parley under silk in the open),
  ℹtem \textbf{Bell Token} (ring Bell Dawn to convert lethal fallout into payable cost),
  ℹtem \textbf{Repair Beads} (pre−pledged seasons of mending).
  \end{itemize}
  
  \textbf{At the table (mechanical hooks):}
  \begin{itemize}
  ℹtem In forest, under canopy, or along a marked furlong: Position +1 once per scene when acting unseen or disengaging.
  ℹtem When a keystone is correctly identified in fiction, treat sabotage or shoring as DV −1.
  ℹtem Spending a Bell Token in a retreat converts one lethal consequence to a named price or timer.
  ℹtem Breaking Vial Courtesy or Mask−Right gives adversaries Position +1 against the cadre until amends are made.
  \end{itemize}
  
  \textbf{Three Lethai−al ge'hai ``builds'' (PC or cohort templates):}
  
  \textbf{1. Lantern−Thief −−− light control & exits}
  \begin{itemize}
  ℹtem \textbf{Edges:} Athletics, Stealth, Notice; Tags: Moon−Mirror, Smoke Net, Night Cord
  ℹtem \textbf{Moves:} Steal the Lamp (shift a scene into dim−−−your terrain), Draw the Eye (create a false exit), Honor Lane (force pursuers to slow or suffer DV +1)
  \end{itemize}
  
  \textbf{2. Keystone−Reader −−− bridges, ferries, gates}
  \begin{itemize}
  ℹtem \textbf{Edges:} Engineering/Lore (structures), Survival (water), Tinkering
  ℹtem \textbf{Moves:} Name the Stone (spot the load point), HEAL or Break (choose repair−year or collapse), Toll−Binder (chalk a temporary tariff that even foes hesitate to cross)
  \end{itemize}
  
  \textbf{3. Molt−Keeper −−− poison/antidote doctrine}
  \begin{itemize}
  ℹtem \textbf{Edges:} Medicine, Alchemy, Insight
  ℹtem \textbf{Moves:} Paired Vials (harm with an offered cure; secures parley), Clean Cut (field amputation as exit), Venom Truce (threat of escalation traded for withdrawal)
  \end{itemize}
  
  ≠wpage
  
  \section*{II. Lethai−thora Ge'hai −−− Vowel & Veil}
  
  \textbf{Doctrine:} Name the room, bind the hour, tilt the future. They win by re−contexting: changing who counts as host/guest, who speaks under lamp, which ledger (Said/Meant) the court will honor, and what it would cost to ignore either.
  
  \textbf{Training:} Living grammar (syntax drift law), mask etiquette, ledger craft (recording Said & Meant together), witness choreography (who stands under the lamp and when), memory palaces, and mark curation (retiring inherited marks before they turn to curses).
  
  \textbf{Court kit:} travel masks with role−ink, ribbon ledgers, name−beads for witness, bell−cord for single−lamp rites, antidote rings, and context knives (ritual blades that cut bonds after prices are named).
  
  \textbf{Operations menu:}
  \begin{itemize}
  ℹtem \textbf{Two−Ledger Trial:} Force both letter and spirit on the table; the losing party must pay in repair years or reputation.
  ℹtem \textbf{Mask Moot:} Declare roles, then seal parley under silk; shunts a lethal standoff into posture, concessions, and timers.
  ℹtem \textbf{Context Turn:} Move a meeting under a single lamp; what was rumor becomes witness. Truths spoken thereafter stick.
  ℹtem \textbf{Molt Design:} Free a trapped official into a new name and role; avoid civil blood by changing the piece, not the board.
  \end{itemize}
  
  \textbf{Strings / Diamonds they carry:}
  \begin{itemize}
  ℹtem \textbf{Clause−Bead} (insert a contingency into a live deal),
  ℹtem \textbf{Name−Bead} (elevate witness under lamp; testimony counts),
  ℹtem \textbf{Weaver's Writ} (bind a multi−party compact across language drift),
  ℹtem \textbf{Shed−Skin Letter} (sanctioned exit from a coercive identity−−−price named, future favor owed).
  \end{itemize}
  
  \textbf{At the table (mechanical hooks):}
  \begin{itemize}
  ℹtem When Mask−Right is established, treat first parley as Controlled→Dominant (Position +1).
  ℹtem Presenting Said & Meant together gives DV −1 to negotiate clauses; hiding one allows GM to spend SB for False Context complications.
  ℹtem A declared under−beam witness turns a soft truth into a binding statement; refusing to stand under lamp shifts Effect −1 for your side until you do.
  \end{itemize}
  
  \textbf{Three Lethai−thora ge'hai ``builds'':}
  
  \textbf{1. Ribbon−Reader −−− contract surgeon}
  \begin{itemize}
  ℹtem \textbf{Edges:} Diplomacy, Law/Customs, Insight
  ℹtem \textbf{Moves:} Read the Stitch (spot hidden clause), Mercy−Knot (turn harm into a payable price), Seal of Seasons (make a concession last one season only)
  \end{itemize}
  
  \textbf{2. Mask−Herald −−− room control}
  \begin{itemize}
  ℹtem \textbf{Edges:} Performance, Command, Etiquette
  ℹtem \textbf{Moves:} Call the Roles (fix who is host/guest), Veil the Blade (threats become symbols; tempers cool), Seat the Witness (choose who stands under lamp)
  \end{itemize}
  
  \textbf{3. Shed−Singer −−− identity exits}
  \begin{itemize}
  ℹtem \textbf{Edges:} Deception, Medicine/Lore (rites), Sway
  ℹtem \textbf{Moves:} Design the Molt (legal/social escape plan with debt), Cure in the Cup (always carry a public antidote), Word−Cut (ritual severance after price named)
  \end{itemize}
  
  ≠wpage
  
  \section*{III. Friction, Rivalry, Integration}
  \begin{itemize}
  ℹtem \textbf{With Ykrul:} Silence furlongs and Bowl & Board share a border. Ge'hai respect Ykrul placement, but price behavior, not routes. Joint treaties tend to work; joint insults tend to last.
  ℹtem \textbf{With Aeler:} Under−beam witness versus mask−right parley creates jurisdictional puzzles. Aeler Spirit Shields prefer lamps; Lethai−thora prefer silk. Ge'hai learn both.
  ℹtem \textbf{With Lethai−ar:} Many ge'hai are also vowed (Weaver/Serpent). Coil and web add sharp edges: cure−paired poisons, mercy−knots, and the taboo against binding without consent.
  \end{itemize}
  
  \section*{IV. Story Heat (drop−in complications)}
  \begin{itemize}
  ℹtem \textbf{Mark Gone Sour.} A cherished ge'hai mark has lost its last living context. It behaves like a curse (insomnia, compulsions). Repair the story or retire the mark.
  ℹtem \textbf{Lamp vs. Veil.} A frontier court demands under−beam witness; the ge'hai require mask−right for safety. Can both thresholds be honored?
  ℹtem \textbf{The Price of Cure.} A tyrant's enforcer begs a molt. The cure exists; naming its price may break a city.
  ℹtem \textbf{Silence Broken.} A silence furlong is violated by foreign scouts; Lethai−al ge'hai must re−stitch the corridor before a war learns to speak there.
  \end{itemize}
  
  \section*{V. Playing Ge'hai (tone & presence)}
  \begin{itemize}
  ℹtem \textbf{Lethai−al:} move like a solved puzzle; every gesture pre−answers a chase. They speak with hands first. Their mercy is letting you leave by the exit they already cut.
  ℹtem \textbf{Lethai−thora:} attend to breath, posture, and lamp distance. They do not ``win arguments''−−−they decide which truths exist in this room for this season.
  \end{itemize}
  
  Ge'hai do not seek glory. They seek rooms that stop killing people and roads that keep their promises. If you meet them at night, listen: the solution is already hanging from the next branch or folded in the next word; your only choice is whether you can live with its price.
  
\end{document}