\documentclass[10pt,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{left=1.2cm,right=1.2cm,top=1.5cm,bottom=1.5cm}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{mdframed}
\usepackage{needspace}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue}

\newcommand{\tag}[1]{\textsc{#1}}
\newcommand{\clk}[1]{\textit{[#1]}}
\newcommand{\stringitem}[1]{\textit{#1}}
\newcommand{\dusana}[1]{\begin{quote}\itshape #1 \\ --- Dusana of the Raven Road\end{quote}}

\titleformat{\chapter}{\Huge\bfseries}{\thechapter}{20pt}{}
\titleformat{\section}{\Large\bfseries}{\thesection}{15pt}{}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8ex}

\newmdenv[linecolor=black,backgroundcolor=gray!5,innertopmargin=8pt,innerbottommargin=8pt,innerrightmargin=8pt,innerleftmargin=8pt]{infobox}

\begin{document}

\title{\Huge\textbf{The Unfinished Cycle}\\ \large Tales of the Threshold, from a Tulkani Fire}
\author{Told by a broken wagon at the edge of the Direwood\\ \small Set down by Dusana of the Raven Road}
\date{Under the lamp, bread and salt --- and a story that keeps the dark at bay}
\maketitle

\tableofcontents

\chapter*{Preface: The Shadow That Lingers}
\addcontentsline{toc}{chapter}{Preface: The Shadow That Lingers}

A cycle of stories is a cycle of fear. Some fears have faces. Some have names. And some --- the oldest kind --- have neither. They are the hollow spaces between heartbeats, the pause before a knock, the silence after a door closes.

My people are the Tulkani. We build no temples. We carry the hearth. But we do not speak of that which listens. Not directly. That which listens. That which watches. That which waits for an invitation.

This autumn, the road was kind. We found a clearing between the Direwood and the Mistland --- a place where the wind did not howl and the ground was dry. We lit a fire. We poured the first cup for the ancestors, the second for the road, the third for the ones we had lost.

The children asked for stories.

The elders warned: "Do not tell the one about the ninth door. Do not tell the one about the bride who tied nine knots. The force is hungry this time of year."

The children pressed. The elders relented. "Tell stories where the shadow is not. Tell stories of people who made mistakes, who opened doors, who forgot to count. The presence does not need to appear. It is already in the room."

So I told these tales. Nine of them. One for each cup we poured into the earth.

The shadow did not come. But the fire burned low. The children did not sleep. Their eyes were wide, and they stayed close to the wagon wheels, and when the wind shifted, they asked, "What was that?"

"Nothing," the elders said. "Nothing at all."

This book is the record of that night. The stories are not mine. They were told to me by a grave-robber in Zakov, by a lighthouse keeper's daughter on the Sidhi coast, by a blind monk in the Pereshi highlands, and by a woman who survived three days in the silence and came back with no shadow.

I have changed names. I have hidden places. I have left out the ninth word, the ninth knock, the ninth step, because some thresholds should not be crossed even in ink.

The fire is cold. The road is south. The presence is patient.

\dusana{At the edge of the Direwood, under a lamp that flickers nine times. The ink is ash and terror.}

\input{copyright}
\chapter{The First Night: The Thing That Knocks Four Times}

\section*{Told by}
A grave-robber in Zakov who had dug up three kings and one child. He spoke with a tremor in his hands, and he would not look at the door.

\section*{The Tale}
There is a rule in Zakov: three knocks is a guest. Four knocks is something else.

A sailor --- call him Jan --- came home late from the docks. The fog was thick. The lamps were lit. He knocked on his door. One. Two. Three.

A pause. Then a fourth knock.

He should not have opened it. He opened it.

His wife stood in the doorway. Her face was his wife's face. Her dress was his wife's dress. But her eyes were the color of the sea at midnight, and her mouth was closed, and she did not blink.

"Come in," Jan said.

She walked past him. She did not take off her coat. She sat in the corner of the room. She did not speak.

Jan asked, "Where have you been?"

The thing that looked like his wife tilted its head. Its neck made a sound like dry wood snapping.

"I have been where the tide goes when it leaves the shore. I have been in the space between your last breath and your next. I have been in the place of unfinished things."

Jan did not sleep that night. He sat across from the thing, watching it watch him. It did not blink. It did not breathe.

At dawn, his real wife opened the door. She was carrying fish from the market. She saw the thing in the corner. She screamed.

The thing stood. It walked to the door. It paused at the threshold. It knocked four times on the frame. Then it left.

Jan's wife asked, "What was that?"

Jan said, "I do not know."

But that night, the thing came back. It knocked four times. Jan did not open the door. The thing knocked again. Four times. Then it began to whisper.

"Let me in. I am still your wife. I have only been gone a little while."

Jan pressed his hand against the wood. The wood was cold. The whispers continued until dawn.

On the third night, Jan opened the door. His wife was there --- his real wife, weeping, saying she had been locked out for three days. He let her in. He held her.

But at midnight, he heard knocking from the bedroom. Four times. He opened the door. The thing was sitting on his bed, wearing his wife's skin. The real wife was gone.

The thing smiled. "You let me in. Now I am here to stay."

Jan lived with the thing for ten years. It cooked his meals. It mended his clothes. It never spoke except to whisper his name when he was alone.

When he died, the thing walked out of the house and knocked on the neighbor's door. Four times.

\section*{The Witch's Closing}
"Three knocks is a guest. Four is an invitation. Do not answer the fourth. Do not open the door. The force does not need to knock twice."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Fourth Knock} --- A spirit that mimics a loved one. It announces itself with four knocks. If the door is opened, the spirit replaces the loved one. The original is lost in the place of unfinished things.

\textbf{Tags / Mechanics:} \tag{SPIRIT} \tag{THRESHOLD} \tag{DREAD}

\begin{itemize}
\item When a character hears four knocks, they must make a \textit{Spirit + Resolve} test (DV 5) to resist opening the door. On failure, the spirit enters. The character's loved one (selected by the GM) vanishes.
\item The spirit will impersonate the loved one perfectly except for its eyes (midnight blue) and its refusal to blink. A character who notices these signs and confronts the spirit may force it to leave by speaking the loved one's true name three times. This costs 1 Corruption.
\item \textbf{Clock: The Hollow's Patience [6]} -- ticks each day the spirit remains in the house. When full, the spirit claims another family member. The only way to stop it is to find the original loved one in the place of unfinished things --- a journey that requires crossing a threshold that does not want to be found.
\end{itemize}

\textbf{Adventure Hook:} A Zakov fisherman has not left his house in a week. His neighbors hear four knocks every night at midnight. The party must enter the house and discover whether the man inside is still human --- and if not, how to drive the thing out before it knocks again.
\end{infobox}

\chapter{The Second Night: The Mother's Needle}

\section*{Told by}
A lighthouse keeper's daughter on the Sidhi coast, who had seen her mother sew a dead man's mouth shut and then pretend it never happened.

\section*{The Tale}
The Sidhi say that a needle can sew more than cloth. It can sew a mouth closed. It can sew a shadow to a wall. It can sew a door shut from the inside.

A seamstress --- call her Yara --- found a needle in her dead mother's sewing box. The needle was black, warm to the touch, and it had no eye. She could not thread it. But when she held it, she could feel it pull toward her mother's grave.

She walked to the cemetery. The needle pulled her to a grave that was not her mother's. The name on the stone was worn away.

"What is this?" she asked.

The needle did not answer. But a voice spoke from the earth. "That is the grave of the woman who taught your mother to sew. She died with a secret in her mouth. Sew it shut, and the secret will stay buried."

Yara did not understand. She took the needle home.

That night, her dead mother sat at the foot of her bed. Her mouth was open. Inside, instead of a tongue, there was a thread of black silk.

"Sew me closed," her mother said. "The force is using my mouth to speak. If you do not sew it shut, I will say your true name, and you will never wake."

Yara tried to sew her mother's mouth. The needle passed through the lips without leaving a mark. The thread would not hold.

"You need the black thread," her mother said. "The thread that was used to sew my shroud. It is in the box."

Yara found the thread. It was the same black as the needle. She sewed her mother's mouth closed. Her mother nodded, stood, and walked through the wall.

The next morning, Yara's own mouth was sewn shut. She could not eat. She could not speak. She could only point.

The village healer cut the threads. They grew back within the hour.

A völva came. She said, "You have taken the needle from that which collects. It has taken your voice in return. The only way to free yourself is to find the original owner of the needle and return it."

"But she is dead," Yara wrote.

"Then find her ghost. And sew her mouth shut again. The first time, you sewed your mother. The second time, you must sew the teacher. The third time, you must sew the one who taught the teacher. Nine generations. Nine mouths. Nine needles."

Yara walked into the silence. She did not come back. But sometimes, on quiet nights, you can hear the sound of a needle piercing cloth --- and a woman whispering, "Sew. Sew. Sew."

\section*{The Witch's Closing}
"A needle with no eye is a door with no handle. Do not pick it up. Do not ask what it sews. The thread of unfinished business is unbreakable, and it is always long enough."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Blind Needle} --- A black needle with no eye. It can sew any material, including flesh, shadow, and memory. Those who use it lose their voice until they sew nine mouths closed. The ninth mouth is always their own.

\textbf{Tags / Mechanics:} \tag{ARTIFACT} \tag{CURSE} \tag{DREAD}

\begin{itemize}
\item A character who picks up the Blind Needle gains the Condition \textit{Silenced} (cannot speak) until they sew nine mouths closed. Each mouth belongs to a dead person whose ghost is trapped in the place of unfinished things.
\item To sew a mouth, the character must find the ghost and succeed on a \textit{Wits + Craft} test (DV 5). On success, the ghost is released and the character may speak one sentence. On failure, the needle sews the character's own mouth partially closed (--1 die to all verbal rolls for a week).
\item \textbf{Clock: The Hollow's Thread [9]} -- ticks each time a mouth is sewn. When full, the character's own mouth is sewn permanently. They can never speak again. The needle vanishes and appears in the hand of a loved one.
\end{itemize}

\textbf{Adventure Hook:} A Sidhi seamstress has stopped speaking. Her family found her holding a black needle, her mouth sewn with thread that cannot be cut. The party must learn who she was trying to silence --- and whether the ninth mouth is worth the price.
\end{infobox}

\chapter{The Third Night: The Guest Who Never Left}

\section*{Told by}
A blind monk in the Pereshi highlands who lived alone in a temple with seven locked doors. He had not opened the seventh door in forty years.

\section*{The Tale}
A traveler came to a village in the monsoon season. He was wet, tired, and polite. He asked for shelter. A widow offered him her spare room.

"Stay one night," she said. "No more."

The traveler agreed. He slept. He ate. He thanked her. In the morning, he left.

But that evening, he returned. "The road is flooded," he said. "May I stay one more night?"

The widow agreed. The traveler stayed. The next morning, he left again. That evening, he returned again.

This continued for a week. The widow grew uneasy. "Why do you keep coming back?"

The traveler smiled. "I never left. You only thought I did. The force lets me step out of your sight, but I am always in the corner of your eye. When you look away, I step closer."

The widow tried to lock the door. The traveler was already inside. She tried to burn the house. The traveler sat in the flames, smiling.

"Wherever you go, I will be there. I am the guest who never leaves. I am the hospitality of unfinished business."

The widow fled. She traveled for a year. She crossed mountains, rivers, deserts. Every night, when she opened her eyes, the traveler was sitting at the foot of her bed.

She returned to the village. She burned her house to the ground. The traveler stood in the ashes.

"I told you," he said. "I never left."

She stopped running. She sat down in the ashes. The traveler sat across from her. They are still sitting there, the widow and the guest, waiting for someone to invite them inside.

\section*{The Witch's Closing}
"Some guests do not need a door. They enter through your eyes. Do not look away from a stranger. The force is patient, and its guests are always polite."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Polite Guest} --- A spirit that attaches itself to anyone who offers shelter to a stranger for more than one night. It cannot be seen directly, only in peripheral vision. It never attacks. It only waits. Its presence drives the victim mad.

\textbf{Tags / Mechanics:} \tag{SPIRIT} \tag{CURSE} \tag{DREAD}

\begin{itemize}
\item A character who offers shelter to a stranger for a second night gains the Condition \textit{Watched}. The Guest appears in their peripheral vision once per day. The character must make a \textit{Spirit + Resolve} test (DV 4) or mark 1 Fatigue (the stress of being watched).
\item The Guest cannot be killed, banished, or reasoned with. It can only be passed on. If the character offers shelter to another stranger and the stranger accepts, the Guest transfers to the new host. The original host is freed.
\item \textbf{Clock: The Guest's Patience [8]} -- ticks each day the Guest remains with a host. When full, the Guest becomes visible. The host sees it clearly for the first time. The sight causes permanent madness (the character loses 1 point of Spirit, minimum 1).
\end{itemize}

\textbf{Adventure Hook:} A Pereshi widow has not slept in three months. She sees a figure in the corner of her vision every time she closes her eyes. The party must find a stranger willing to take the Guest --- or discover another way to break the hospitality of unfinished business.
\end{infobox}

\chapter{The Fourth Night: The Hand Under the Bed}

\section*{Told by}
A child who survived a night in the Direwood. The child was not afraid of the dark. The child was afraid of what the dark touched.

\section*{The Tale}
A little boy --- call him Tomas --- was afraid of the hand under his bed. Every night, he felt it brush against his ankle. Every morning, he looked. Nothing was there.

His mother said, "It is your imagination."

His father said, "There is nothing under your bed."

Tomas asked the village priest. The priest said, "That which listens does not hide under beds. It hides in the space between the floor and the ground. That space is very small. But it can fit anywhere."

That night, Tomas did not sleep. He lay still, feeling the hand. It was cold. It had too many fingers.

He whispered, "What do you want?"

The hand stopped. A voice came from under the bed. It was the voice of a child who had been dead for a hundred years.

"I want you to look."

Tomas looked. The hand was attached to an arm, and the arm was attached to a body, and the body was folded into the space under the bed like a letter folded into an envelope. The body had no face. Where the face should have been, there was a second hand.

Tomas screamed. His parents ran in. They looked under the bed. Nothing was there.

But Tomas had seen. And seeing meant it had seen him back.

For the rest of his life, Tomas felt the hand under his bed. Even when he grew up and moved away. Even when he slept in a house with no bed, only a mat on the floor. The hand was there. It brushed his ankle. It waited.

One night, Tomas reached down. He touched the hand. The hand pulled him under the bed. He folded himself into the space, smaller and smaller, until he fit.

His body was found the next morning, tucked under the bed, folded like a letter. His face was gone. In its place was a second hand.

\section*{The Witch's Closing}
"Do not look. Do not touch. The hand is patient. It will wait until you are small enough to fit."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Hand Under the Bed} --- A spirit that lives in the space between the floor and the ground. It touches sleepers' ankles. If the sleeper looks, the spirit shows its face --- which is a hand. The sleeper is then marked for death.

\textbf{Tags / Mechanics:} \tag{SPIRIT} \tag{DREAD} \tag{THRESHOLD}

\begin{itemize}
\item A character who looks under their bed after feeling a touch must make a \textit{Spirit + Resolve} test (DV 6). On failure, the character gains the Condition \textit{Marked}. The Hand will touch them every night. Each night, the character marks 1 Fatigue.
\item After seven nights, the character must make another test (DV 8). On failure, the Hand pulls them into the space under the bed. The character disappears. They cannot be recovered by any means short of a major ritual that requires a sacrifice of equal weight (another Marked victim).
\item The Hand cannot be killed. It can be avoided by sleeping on the ground floor (no space under the bed) or by sleeping in a hammock. The Hollow does not like hammocks.
\end{itemize}

\textbf{Adventure Hook:} A village has lost seven children in seven months. Each child was found under their bed, folded into a space too small for a body. The party must discover what the children saw --- and whether any of them survived the looking.
\end{infobox}

\chapter{The Fifth Night: The Road That Eats Names}

\section*{Told by}
A woman who survived three days in the silence and came back with no shadow. She spoke from behind a curtain, and her voice was the sound of wind through dead grass.

\section*{The Tale}
There is a road that does not appear on any map. It runs between the Direwood and the Mistland, and it has no beginning and no end. The Linn call it the Name-Eater.

A traveler --- call her Mira --- took the Name-Eater by accident. She was trying to reach the coast. The fog rolled in. The road under her feet changed from dirt to something smooth and gray, like stone worn by water.

She walked for an hour. The fog cleared. She was in a place she did not recognize. The trees were silver. The sky was the color of a bruise.

A figure stood in the road. It had no face, only a mouth. The mouth opened.

"Give me your name."

Mira refused. She tried to walk around the figure. The road curved, and the figure was there again.

"Give me your name."

Mira ran. The figure did not follow. But the road began to whisper. It whispered names. Names of people Mira had known. Names of people she had forgotten. Names of people who had died before she was born.

"Give me your name. Everyone else has. You are the last."

Mira looked behind her. The road was littered with coins, shoes, hairpins, and teeth. Each one, the figure said, was a name. Someone had given their name and turned into that object.

Mira took a tooth from the road. She put it in her mouth. She bit down.

The figure screamed. "That is not your name. That is the name of a dead man. You cannot give what is not yours."

Mira walked past the figure. She walked for three days. When she emerged from the road, she was in a village she had never seen. The people asked her name. She opened her mouth to speak, but no sound came out.

She had not given her name. But she had held a dead man's name in her mouth. The dead man had not given it to her. She had stolen it. And the force that governs debts considered theft as good as payment.

She lived the rest of her life without a name. She pointed. She wrote. But no one could remember her from one day to the next. She was a stranger even to herself.

\section*{The Witch's Closing}
"The Name-Eater does not need your consent. It only needs you to hold a name that is not yours. Do not pick up teeth on the road. Do not answer when the road whispers."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Name-Eater} --- A road that appears in fog between the Direwood and the Mistland. It is lined with objects (coins, teeth, hairpins) that are the transformed names of those who walked before. To escape, a traveler must give a name --- either their own or someone else's. Giving another's name works, but the other person becomes nameless.

\textbf{Tags / Mechanics:} \tag{LOCATION} \tag{CURSE} \tag{DREAD}

\begin{itemize}
\item A character who walks the Name-Eater must make a \textit{Wits + Resolve} test (DV 5) every hour to avoid picking up an object from the road. On failure, they pick up an object and gain the Condition \textit{Nameless} (cannot speak their own name, and others forget them after one day).
\item To leave the road, the character must offer a name to the figure. They may offer their own name (permanent loss of name, as above) or the name of another person they know. Offering another's name causes that person to become nameless instead. The character is freed.
\item \textbf{Clock: The Road's Hunger [8]} -- ticks each time a name is taken. When full, the road swallows the nearest village. Everyone in the village loses their names. The village disappears from all maps.
\end{itemize}

\textbf{Adventure Hook:} A village has vanished from the edge of the Direwood. The only trace is a road of gray stone leading into the fog. The party must walk the Name-Eater to find the village --- but every step costs a name, and they have brought only their own.
\end{infobox}

\chapter{The Sixth Night: The Song the Dead Sing to Calves}

\section*{Told by}
A Tulkani herder who lost a third of his herd to a winter sickness. He swore the calves had heard something before they died.

\section*{The Tale}
The Tulkani say that the dead sing to calves. Not to cows. Not to bulls. Only to the young, the innocent, the ones who have not yet learned to fear.

A herder --- call him Ivan --- woke to find his calves standing in a circle. They were not sleeping. They were listening. Their ears were turned toward the ground.

"What do you hear?" Ivan asked.

The calves did not answer. They could not. They were calves.

Ivan put his ear to the ground. He heard singing. It was not in any language he knew. It was the sound of water freezing, of ice cracking, of a door closing.

He pulled away. His ear was cold. The skin was black with frostbite.

That night, the calves lay down and did not rise. Their eyes were open. Their mouths were open. A thin thread of ice extended from each mouth into the earth.

The völva came. She touched the ice. "The force is singing to your calves. It is using their breath to freeze the ground. When the ground is frozen deep enough, the dead will walk."

"How do I stop it?" Ivan asked.

"You cannot. The calves are already dead. They just do not know it yet. You must burn the bodies before the song ends."

Ivan burned the calves. The ice melted. The ground thawed. But that night, he heard singing from his own bed. His daughter was sitting up, her mouth open, a thin thread of ice extending from her lips.

He put his hand over her mouth. The ice cut him. He wrapped her in a blanket and held her until dawn. The singing stopped.

His daughter never spoke again. But she never froze either. She grew up with a scar on her tongue where the ice had been.

\section*{The Witch's Closing}
"The dead sing to the young because the young still remember how to listen. Cover your children's ears when the ground is cold. The lullaby is sweet, and the sleep it offers is eternal."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Frozen Lullaby} --- A song sung by the dead that rises from the ground on winter nights. It affects the young (children and animals) first. Those who hear it freeze from the inside out. The ice can be melted with fire, but the listener may lose their voice or their hearing.

\textbf{Tags / Mechanics:} \tag{PHENOMENON} \tag{DREAD} \tag{THRESHOLD}

\begin{itemize}
\item When the temperature drops below freezing, the GM may declare that the Hollow is singing. All children (NPCs) and young animals within earshot must make a \textit{Spirit + Resolve} test (DV 5) or begin freezing. The character has one hour to warm them (fire, blankets, body heat) or they die.
\item Adult characters who hear the song must also test (DV 4) or gain the Condition \textit{Frost-Tongue} (cannot speak for 1d4 days; the vocal cords are damaged). Permanent loss of voice is possible on a critical failure.
\item The song can be blocked by covering the ears with wax or cloth, but the song vibrates through the ground. Characters standing on bare earth cannot block it completely.
\end{itemize}

\textbf{Adventure Hook:} A Tulkani caravan has stopped in a valley where the ground never thaws. The children have stopped speaking. The calves are dead. The herders say they hear singing every night. The party must find the source of the song --- and silence it before the next freeze.
\end{infobox}

\chapter{The Seventh Night: The Bell That Rings Backwards}

\section*{Told by}
An Aelinnel bell-tender who had lost three fingers to a frozen clapper. He spoke in a whisper, and his hands were wrapped in gray wool.

\section*{The Tale}
Bells are for the living. They ring forward, through time, from the strike to the echo. But there are bells that ring backwards.

A gnome --- call him Pim --- found a bell in the ruins of a phasing city. The bell was small, bronze, and warm. He rang it once. The sound was wrong. It went from silence to ring, then from ring to silence, but the silence came before the ring.

"It is ringing backwards," his wife said. "Throw it away."

Pim did not throw it away. He hung it in his workshop. Every day, he rang it once. The sound made him feel strange. He forgot things. He forgot his wife's name. He forgot the way home.

One day, he rang the bell, and his wife disappeared. Not suddenly. Slowly. She was there, and then she was not. He remembered her fading, like a reflection in a pool that someone had stirred.

The bell was cold now. He rang it again. His workshop disappeared. He was standing in the street, holding the bell, wearing clothes he did not recognize.

He rang it a third time. He forgot why he was holding the bell. He dropped it. It hit the ground and rang on its own. The ring was forward this time. It was a normal bell sound.

Pim remembered his wife. He remembered his workshop. He remembered his name. But he could not find the bell. It had rolled into a crack in the pavement.

He searched for years. He never found it. But sometimes, late at night, he hears a bell ringing backwards. He knows it is the bell of unfinished business, calling him home.

\section*{The Witch's Closing}
"A bell that rings backwards is a door that opens into yesterday. Do not ring it. Do not listen. Its time is not our time, and its memory is longer than yours."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Backward Bell} --- A bronze bell that rings in reverse. Each ring erases a memory (starting with recent events, then names, then places). After three rings, the ringer forgets everything. The bell then rings forward, restoring all memories --- but the bell disappears. It will reappear in another location, waiting for a new victim.

\textbf{Tags / Mechanics:} \tag{ARTIFACT} \tag{CURSE} \tag{DREAD}

\begin{itemize}
\item Ringing the bell once erases one recent memory (GM's choice, minor). The character may spend 1 Boon to resist.
\item Ringing the bell twice erases a significant memory (a name, a face, a skill). The character loses 1 XP (the memory is gone; the XP must be re-earned).
\item Ringing the bell three times erases the character's identity. They become a blank slate. The bell then rings forward, restoring all memories --- but the character must then find the bell again to regain what was lost. The bell will have moved.
\item \textbf{Clock: The Hollow's Echo [4]} -- ticks each time the bell is rung. When full, the bell rings itself. Everyone within earshot loses one memory.
\end{itemize}

\textbf{Adventure Hook:} A gnome inventor has gone missing. His workshop is empty except for a hook on the wall where a bell once hung. The party must find the bell before it rings again --- and before they forget why they were looking.
\end{infobox}

\chapter{The Eighth Night: The Guest Who Left a Shadow}

\section*{Told by}
A Sidhi dock-worker who had once found a shadow pinned to a wall with a nail. The shadow still moved when the wind blew.

\section*{The Tale}
A woman --- call her Sasha --- opened her door to a stranger. The stranger was thin, pale, and wore a coat made of gray wool. He asked for water. She gave it. He asked for bread. She gave it. He asked for a place to sit.

Sit by the fire," she said.

The stranger sat. He did not warm his hands. He did not speak. He stared at the fire until it went out.

When Sasha woke the next morning, the stranger was gone. But on the wall where he had sat, there was a shadow. It was not her shadow. It was his.

The shadow moved. It gestured. It pointed at the door.

Sasha tried to ignore it. But the shadow grew larger. It reached out and touched her own shadow. Her shadow flinched.

She called a priest. The priest said, "The presence has left its mark on your wall. The shadow is a door. If you let it grow, the stranger will return."

"How do I stop it?"

"You must nail the shadow to the wall. Use an iron nail. Strike it through the shadow's heart."

Sasha took a nail. She hammered it into the wall where the shadow's chest would be. The shadow screamed. It twisted. It pulled at the nail. But it could not escape.

The stranger returned that night. He stood outside her door, knocking. "Let me in. I need my shadow."

"Your shadow is mine now," Sasha said. "I paid for it with iron."

The stranger laughed. "Iron does not belong to you. It belongs to the earth. And the earth belongs to the force that collects."

He walked through the door. He pulled the nail from the wall. The shadow wrapped around him like a cloak. He smiled.

"Thank you for keeping it warm."

He left. Sasha's own shadow was gone. She lived the rest of her life without one. In the sun, she looked like a hole in the world.

\section*{The Witch's Closing}
"Do not nail a shadow to a wall. You will only make it angry. The force does not forget its possessions. And a shadow is always its, no matter how long you keep it."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Nailed Shadow} --- A shadow left behind by a Hollow traveler. It can be pinned with an iron nail, but the nail only holds the shadow for one night. The next night, the traveler returns to reclaim it. The person who nailed the shadow loses their own shadow forever.

\textbf{Tags / Mechanics:} \tag{SPIRIT} \tag{CURSE} \tag{DREAD}

\begin{itemize}
\item A character who nails a Hollow shadow to a wall gains the Condition \textit{Shadowless}. They cast no shadow. This gives --1 die on Stealth rolls (no shadow to hide behind) and +1 die on social rolls involving fear (others are unsettled).
\item The Hollow traveler will return at midnight to reclaim the shadow. The character may choose to fight the traveler (impossible to kill, but can be banished with a successful \textit{Spirit + Lore} test, DV 7). On failure, the traveler takes the character's own shadow as interest.
\item \textbf{Clock: The Shadow's Memory [4]} -- ticks each day the character is shadowless. When full, the character forgets what shadows are. They cannot recognize reflections, silhouettes, or any two-dimensional representation of a living thing.
\end{itemize}

\textbf{Adventure Hook:} A Sidhi innkeeper has a shadow nailed to her wall. She does not know whose it is, but every night at midnight, something knocks on her door. The party must decide: remove the nail and let the shadow go, or find a way to destroy it before the traveler returns.
\end{infobox}

\chapter{The Ninth Night: The Door That Opens Both Ways}

\section*{Told by}
The fire itself. Or so we dreamed. The flames were low, and the shadows were long, and when we woke, the story was already in our mouths.

\section*{The Tale}
There is a door in the Direwood that opens both ways. It is not a door of wood or stone. It is a door of air, of silence, of the space between two breaths.

A hunter --- call him Leif --- found the door while chasing a white stag. The stag ran through the door. Leif followed.

On the other side was the place of unfinished things. The sky was gray. The ground was gray. There were no trees, no birds, no wind. Only a path of gray stones leading into the distance.

Leif walked for days. He did not get hungry. He did not get tired. He did not get thirsty. He just walked.

He found the white stag standing at the end of the path. The stag turned to look at him. Its eyes were the color of the sky.

"You are not supposed to be here," the stag said.

"Then let me leave," Leif said.

"I cannot. The door you came through only opens one way from your side. From here, it opens into a different place. You will not recognize it. You will not be welcome."

Leif did not care. He walked back to the door. He stepped through.

He was in his village. But the village was wrong. The houses were the same, but the people had different faces. His wife did not know him. His children ran away.

He had been gone for a hundred years. The door had opened into the future.

Leif lived as a stranger in his own home. He died alone, and the villagers buried him at the edge of the woods. They did not mark the grave.

The white stag still runs through the Direwood. And the door still opens both ways. If you find it, do not follow. The force does not care which direction you go. It only cares that you go.

\section*{The Witch's Closing}
"A door that opens both ways is a door that has no inside and no outside. Step through, and you will never be sure which side is home. The force does not lie. It simply does not tell you which way the door is facing."

\section*{What Lurks in the Shadow of the Tale}
\begin{infobox}
\textbf{The Two-Way Door} --- A threshold in the Direwood that leads to the Hollow. Stepping through from the mortal side leads to the Hollow. Stepping back from the Hollow leads to a random time and place (GM's choice). The door does not guarantee return to the same moment.

\textbf{Tags / Mechanics:} \tag{LOCATION} \tag{THRESHOLD} \tag{DREAD}

\begin{itemize}
\item A character who steps through the Two-Way Door must make a \textit{Wits + Lore} test (DV 6) to understand where they are. On failure, they are lost in the Hollow for 1d10 days (mark 1 Fatigue per day). They emerge at a random location and time.
\item Stepping back through the door from the Hollow side always leads to a different place and time than the one the character left. The GM chooses: up to 100 years forward or backward, or up to 1000 miles away.
\item The door cannot be destroyed. It can be sealed with a ritual that requires the blood of a white stag and a promise never to hunt in the Direwood again. Breaking the promise unseals the door.
\end{itemize}

\textbf{Adventure Hook:} A hunter has been missing for three days. His wife found his tracks leading into the Direwood. The party must follow the tracks --- but the tracks are a hundred years old, and the hunter's grandson is the one who hired them.
\end{infobox}

\chapter*{Epilogue: The Fire That Listened}
\addcontentsline{toc}{chapter}{Epilogue: The Fire That Listened}

The stories ended. The fire was low. The children were asleep, though some of them were pretending. The elders stirred the coals.

"Did it come?" a child whispered.

"No," an elder said. "It does not come when you tell stories about it. It comes when you forget to tell them."

The wind shifted. A branch tapped against the wagon. Four times.

We did not look. We did not count. We added wood to the fire and spoke of other things. The road. The weather. The price of salt.

At dawn, we left the clearing. The ground where the fire had been was bare. No ash. No coals. Only a patch of black earth, warm to the touch.

We did not thank it for its hospitality. We did not acknowledge that it had been there, listening, learning our names.

The ninth cup was poured into the earth. We did not look to see who drank it.

\dusana{On the road south, with a notebook full of fear and a fire that still burns in memory. The presence is patient. The stories are told. The threshold holds --- for now.}

\chapter*{Appendix: Signs of the Unfinished}
\addcontentsline{toc}{chapter}{Appendix: Signs of the Unfinished}

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{p{4cm} X}
\toprule
\textbf{Sign} & \textbf{What It Means} \\
\midrule
Four knocks at night & The presence is at the door. Do not answer. \\
A needle with no eye & It sews mouths closed. Do not touch. \\
A guest who never leaves & Its hospitality. Do not offer a second night. \\
A hand under the bed & It is reaching. Do not look. \\
A road of gray stone & The Name-Eater. Do not pick up teeth. \\
Singing from the ground & Its lullaby. Cover your ears. \\
A bell that rings backwards & Its time. Do not ring it. \\
A shadow pinned to a wall & Its mark. Do not nail it. \\
A door in the Direwood & Its threshold. Do not step through. \\
\bottomrule
\end{tabularx}
\end{table}

These are not warnings. They are courtesies extended to that which listens. Break them, and it will not punish you. It will simply invite you in.

And its hospitality is eternal.

\chapter*{Colophon}
\addcontentsline{toc}{chapter}{Colophon}

This book was written on paper made from linen rags, in ink cut with ash and the frost of a winter night. The binding is leather, tanned with the hide of a calf that heard the song.

The ninth page is blank.

The ninth story is the one we did not tell.

It listened.

\dusana{At the edge of the Direwood, with a notebook full of fear. The ink is fresh. The fire is cold. The threshold is waiting.}
\end{document}