\documentclass[11pt]{book}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a5paper, margin=1.2in}
\usepackage{graphicx}
\usepackage{epigraph}
\usepackage{lettrine}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{fontspec}
\setmainfont{Linux Libertine O}
\setsansfont{Linux Biolinum O}
\setmonofont{Inconsolata}

\titleformat{\chapter}{\huge\bfseries\scshape}{\thechapter}{20pt}{}
\titleformat{\section}{\Large\bfseries}{\thesection}{15pt}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{12pt}{}

\defineauthor{gray}{The Grey Wanderer}
\defineauthor{valerius}{Lord Valerius Valvano}
\defineauthor{aqyl}{Aqyl, son of Aqyl}
\defineauthor{survivor}{A Retainer}
\defineauthor{heir}{The Young Heir}

\newcommand{\sb}[1]{\textbf{SB #1}}
\newcommand{\dv}[1]{\textbf{DV #1}}
\newcommand{\clock}[1]{\textbf{[#1]}}

\begin{document}

\frontmatter
\title{\Huge\scshape The Exile Road\\\large A Fenwood Legacy Adventure}
\author{Found in the archives of Tarlington, copied by a faithful hand}
\date{Year of the Fourth Sack, 876 AR}
\maketitle

\tableofcontents

\mainmatter

\chapter*{A Note on This Writing}
\addcontentsline{toc}{chapter}{A Note on This Writing}
\lettrine{T}{his} ledger was found in a sealed chest beneath the Fenwood keep, wrapped in oilcloth and tied with a braid of Ykrul horsehair. The pages are foxed, some scorched, others blurred by water from the fens. Three hands appear: one aristocratic and proud, the second a cramped scholar's script, the third a child's scrawl overwritten by a mother's corrections. A fourth hand---perhaps the archivist---has added marginal notes in a drier ink.

We have transcribed the leaves as they were found, preserving gaps, lacunae, and the occasional whiff of bog-mist. The narrators speak in their own voices, through their own wounds. Let the reader beware: this is not a chronicle of triumph. It is a record of what a house must bury to survive.

\vspace{1em}
\noindent\textit{--- The Grey Wanderer, at a caravanserai in the Violet Steppe, spring of 876 AR}

\clearpage

\chapter{Introduction: The Third Sack}

\lettrine{T}{he} Valvano dynasty is dead. The Third Sack has ended our line---or so the heralds of the usurper proclaim. What remains is a cadet branch: a handful of cousins, servants, heirlooms, and a name that now draws pursuit like a wounded hart draws wolves.

This is the tale of our flight. It is written for our descendants, should any survive, and for the ledger of the Everflame, which alone witnessed our oaths.

\section*{What You Need}
\textbf{Tier:} Seasoned (41--90 XP) \quad \textbf{Players:} 3--6 \quad \textbf{Length:} 3--4 sessions

\subsection*{Core Clocks}
\begin{itemize}
\item \textbf{Valvano Loyalty [6]} --- Clinging to the old name and claims. Advances when you refuse compromise, perform imperial rites, or reject "barbarian" customs. When full, you cannot abandon the name.
\item \textbf{Survival [8]} --- Supplies, allies, shelter. Advances when you lose provisions, suffer setbacks, or make desperate choices. When empty, someone starves or a heirloom is lost forever.
\item \textbf{Tarling Favor [6]} --- The goodwill of the Viterran house you seek. Advances when you accept humiliation, convert to the Temple of Light, or offer the Laurel Seal as tribute. When full, you are granted a pithy county---the future Fenwood.
\end{itemize}

\subsection*{Heirlooms}
Each heirloom is a Minor Asset (4 XP). Narrate its recovery in play.
\begin{itemize}
\item \textbf{The Laurel Seal} --- A bronze disc stamped with the Valvano crest. Grants \textbf{+1 die} to social rolls when presented (once per scene).
\item \textbf{The Bell Shard} --- A fragment of the Noon Bell that once marked imperial time. When rung, all allied rolls are \textbf{Dominant} for one exchange---but the usurper's agents hear it.
\item \textbf{The Debt Ledger} --- A leather-bound book recording every oath the Valvano dynasty swore. Once per session, you may ask the GM: \textit{"What debt can we call in here?"} Gain a concrete advantage, but incur \textbf{Obligation 1} to the original debtor's heir.
\end{itemize}

\subsection*{A Word on Tone}
This is not a heroic flight. It is a retreat---through burning cities, hostile courts, and fens that swallow hope. Mercy is rare; cruelty is casual. Play not for glory, but for the survival of a memory. And remember: \textit{the road does not forgive a debt. It only reschedules it.}

\clearpage

\chapter{Act I: The Sack --- By Lord Valerius Valvano}

\lettrine{I}{} write this in a chamber that still reeks of smoke. The elder branch is dead. My cousin the emperor fell at the Iron Gate, his body not recovered. My sister threw herself from the Laurel Tower. My sons---I have no sons. Only nephews, and a daughter who will not look at me.

The usurper's wolves are at the walls. We have perhaps one candlemark before the Pilgrim's Gate is sealed.

\section*{Opening Scene: The Burning Manor}
The PCs begin in the Valvano manor house, flames visible through the colonnade. They have minutes to gather what they can. Read aloud:

\begin{quote}
\emph{The smoke stings your eyes. Below, the marble forum has become a charnel pit. A soldier in black plate kicks aside a fallen statue of the Noon Bell. He drinks from a stolen amphora.}

\emph{Aqyl---the old scholar---appears at the stair. He does not run. He has never run. "Choose what you carry," he says. "The ledger. The seal. The bell. Then the gate. I will meet you on the road."}

\emph{Behind you, the chapel's sacred flame gutters. Someone has kicked the brazier. The coal still glows.}
\end{quote}

\subsection*{Recovering the Heirlooms}
Each PC may attempt one recovery roll before the gate seals. \textbf{DV 3}. Success secures an heirloom. \textbf{Partial} secures it but generates \textbf{1 SB} (pursuit, injury, or loss of another item). \textbf{Miss} --- the heirloom is lost, and the PC marks \textbf{Fatigue 1}.

\subsection*{Clock: City Collapse [6]}
Advances by 1 each time a PC fails a recovery roll or when the GM spends \textbf{2 SB} (a building collapses, a patrol arrives). When full, the Pilgrim's Gate slams shut. Any PC still inside must escape through the sewers---\textbf{Desperate Position}, \textbf{DV 3}---or be captured.

\subsection*{The Sacred Coal}
The Everflame chapel's brazier still holds embers. A PC may take a coal (free action, no roll). Doing so binds them to the old faith. They gain \textbf{+1 die} on any social roll involving Everflame clergy in the future---but the Temple of Light will consider them heretics.

\subsection*{Key Choices}
\begin{itemize}
\item \textbf{The Censor's Daughter.} A young woman (Livia) begs to join the flight. She carries a redacted scroll from the Hall of Censors. Trust her? She may be a spy for the Witness cult---or its last survivor.
\item \textbf{The Ledger.} The debt ledger is in a burning scriptorium. To save it, a PC must enter a collapsing room (\textbf{Body + Athletics}, \textbf{DV 3}). On a miss, they take \textbf{Harm 1} but retrieve the book.
\item \textbf{The Oath-Brother.} A Ykrul mercenary (Rekedim's ancestor?) offers his sword for safe passage to the steppe. He does not kneel. He waits.
\end{itemize}

\subsection*{Spending SB in Act I}
\begin{itemize}
    \item \textbf{1 SB:} A falling beam separates the party; one PC must find another exit.
    \item \textbf{2 SB:} The usurper's herald recognizes the Laurel Seal; pursuit gains \textbf{+1 die}.
    \item \textbf{3 SB:} The sacred coal ignites a tapestry; \textbf{Environmental Hazard} --- all rolls \textbf{Desperate} until extinguished.
    \item \textbf{4+ SB:} A Valvano elder is captured. The PCs hear his screams as they flee.
\end{itemize}

\section*{The Pilgrim's Gate}
When the City Collapse clock reaches 6 (or when the PCs choose to flee), they reach the gate. Aqyl is already there, holding a lantern. Behind them, the manor collapses.

Read:
\begin{quote}
\emph{The old scholar does not look back. "Names are thresholds," he says. "Some you cross. Some you burn. Tonight, we cross." He steps into the fog. The lantern's flame burns blue.}
\end{quote}

\subsection*{Act I Resolution}
Advance \textbf{Survival} by the number of heirlooms lost (none lost = --2). Advance \textbf{Valvano Loyalty} by the number of imperial rites performed (taking the coal counts as +1). If the Ykrul mercenary joined, note \textbf{The Braid Token}---a future ally.

\clearpage

\chapter{Act II: The Road --- By Aqyl, son of Aqyl}

\lettrine{I}{} am not a Valvano. I am older. The first Aqyl compiled the Corpus Canr\'e when the empire was young. I have been called his son, his father, his shade. The truth is less interesting. I keep the witness.

The road east is a ledger of failures. Acasia frays like old rope. Silkstrand---then called Ceylonne---welcomes coin, not refugees. We are pursued by rumor, not men. That is worse.

\section*{The Ykrul Ford}
The party reaches a river crossing held by the Grey Ash clan. Their chieftain, Khatun Sarnai, blocks the ford.

Read:
\begin{quote}
\emph{Sarnai does not rise from her felt stool. "I know that seal," she says, pointing at the Laurel. "Your great-grandmother gave it to my grandmother as a hostage gift. She left us a son. The braid still hangs in my tent."}

\emph{The Ykrul riders do not draw steel. They wait.}
\end{quote}

\subsection*{The Price}
Sarnai offers passage---for a price. The party may choose one:
\begin{enumerate}
    \item \textbf{Acknowledge the mixed-heritage child.} Publicly accept the half-Ykrul offspring as Valvano. \textbf{Valvano Loyalty --2}, \textbf{Survival +1} (safe passage), gain \textbf{The Braid Token} (future Ykrul favor).
    \item \textbf{The Duel.} A champion fights Sarnai's best rider. \textbf{Body + Melee}, \textbf{DV 4}. Win: passage, no loss. Lose: surrender the Laurel Seal (\textbf{Valvano Loyalty --1}) or mark \textbf{Harm 2}.
    \item \textbf{The Steppe Oath.} Swear a future alliance. Gain passage now, but \textbf{Obligation [4]} to the Grey Ash---they will call in the debt during the fourth sack.
\end{enumerate}

\subsection*{The Child}
If the party acknowledges the mixed blood, Sarnai produces a youth---half-Ykrul, half-Valvano---who has been raised among the clan. He has no name. The party may take him (gaining \textbf{The Braid Token} and a future follower) or leave him (no cost, but \textbf{Valvano Loyalty +1} for purity). Aqyl watches, silent.

\subsection*{Clock: Supplies [6]}
The Ykrul ford delays the party. Each failed negotiation advances the Supplies clock by 1. When it reaches 6, the party must barter a heirloom or mark \textbf{Fatigue 2} for all.

\subsection*{The Linnic Mercenaries}
At a crossroads, a band of sea-raiders (ancestors of the Everbloods) offers escort for a share of future lands. Their captain wears a wolf"s-tail cloak.

Read:
\begin{quote}
\emph{"The Tarlings are dying," the captain says. "Help us take their country, and we will make you counts. The old kings are ghosts. We offer flesh and steel."}
\end{quote}

Aqyl, in a marginal note: \emph{"This is the seed. Remember it. The Everbloods will rise, and the Fenwoods will choose."}

\subsection*{Act II Resolution}
Advance \textbf{Survival} by the number of heirlooms traded (none = --1). If the mixed-heritage child joined, start the \textbf{Unnamed Child} follower track. If the Linnic offer was accepted, note \textbf{Everblood Promise}---a future alliance.

\clearpage

\chapter{Act III: The Rejection --- By a Retainer}

\lettrine{M}{y} lord Valerius is not well. The smoke scarred his lungs. The stares of the Vhasian court scarred his pride. We crossed the border at dawn, expecting sanctuary from House Terinos. They gave us bread and salt---and a message: \textit{"The Tarlings know you are coming. They are curious to meet the grandchildren of their conquerors."}

I am a soldier. I do not trust curiosity.

\section*{The Mocking Feast}
The Vhasian duke (a cousin of the Terinos) invites the Valvanos to a feast. The courtiers wear masks. The wine is sour.

Read:
\begin{quote}
\emph{The duke claps his hands. "Let the exiles prove their lineage! A game of Canray, perhaps? Or a duel? Or---I know---recite the Oath of the Noon Bell. From memory." He smiles. "Your ancestors made us kneel. Now you kneel for entertainment."}
\end{quote}

\subsection*{Trials}
The party must choose a champion and pass two of three trials. Each failure advances \textbf{Patience of the Gate [4]}.
\begin{itemize}
    \item \textbf{Kon'reh.} Play a quick match (use the \textit{Kon'reh Quickstart} or roll \textbf{Wits + Lore}, \textbf{DV 4}). Win: \textbf{Tarling Favor +1}.
    \item \textbf{Duel.} \textbf{Body + Melee}, \textbf{DV 4}. Win: the court applauds; \textbf{Tarling Favor +1}.
    \item \textbf{Oath-Recitation.} \textbf{Presence + Lore}, \textbf{DV 3}. Recite the Oath of the Noon Bell without error. Gain \textbf{+1 die} on all future Valvano Loyalty rolls.
\end{itemize}

If the party refuses all trials, they are expelled. Advance \textbf{Patience of the Gate} by 2.

\subsection*{The Cadet's Betrayal}
A cousin (NPC) offers to sell the Valvano claim to the usurper for safe passage. A PC may detect the plot with \textbf{Wits + Insight}, \textbf{DV 3}. If exposed, the party may execute the traitor (\textbf{Valvano Loyalty --1}, \textbf{Tarling Favor +1}) or exile him (no change). If not exposed, the usurper's agents ambush the party on the road (\textbf{+2 SB} banked for the GM).

\subsection*{The Tarling Messenger}
A rider from Viterra arrives with a sealed scroll.

Read:
\begin{quote}
\emph{"Our grandfathers surrendered to your grandfathers," the messenger reads, his voice flat. "Now you come to our doorstep, asking for crumbs. The gods have a sense of humor. The Tarling duke grants you passage to the fens---on these terms."}
\end{quote}

\textbf{Terms:}
\begin{itemize}
    \item Walk the last hundred miles barefoot (\textbf{Survival --2}, \textbf{Tarling Favor +2}).
    \item Present the Laurel Seal as a gift, not a claim (\textbf{Valvano Loyalty --2}, \textbf{Tarling Favor +2}).
    \item Convert to the Temple of Light (\textbf{Valvano Loyalty --4}, \textbf{Tarling Favor +4}, but lose the Everflame coal's benefit).
\end{itemize}

The party may negotiate (each term reduced by 1 step with a successful \textbf{Presence + Sway}, \textbf{DV 4}) or refuse.

\section*{Act III Resolution}
If \textbf{Patience of the Gate} fills, the Terinos expel the party. They must travel through bandit country (\textbf{Survival --4}) or seek the Ykrul again. Record final \textbf{Tarling Favor}.

\clearpage

\chapter{Act IV: The Fens --- By the Young Heir}

\lettrine{M}{y} name is not yet Fenwood. My father says we must earn the right to bury our name. I do not understand. The fens smell of old death. The tower we are granted is half-collapsed. The local priest of the Temple of Light spat at the ground when he saw my mother's Everflame pendant.

The Tarling duke gave us Ash-Fenn. I asked why it was called that. No one answers.

\section*{Ash-Fenn}
The cursed site is a clearing of blackened earth, surrounded by bog. A single stone marks where a caravan burned.

Read:
\emph{The ground is warm, even in autumn. If you press your ear to the stone, you can hear a chanting---not words, but a thrum, like a bell rung underwater. Aqyl stands apart. He will not approach the stone.}

Aqyl, if asked: \emph{"This is where they burned the Witness cult. The ash fell like rain. The road kept no prints. I was here. I remember the thrum."}

\subsection*{Clock: Winter Provisions [8]}
The party must survive until spring. Each week, roll \textbf{Survival + Wits} (\textbf{DV 3}). Failure advances the clock by 1. When it reaches 8, starvation claims someone (\textbf{PC chooses which NPC follower dies}).

\subsection*{The Bog Witch}
A hedge-witch (a Daughter of the Chain? Morag herself?) appears at the edge of the camp.

Read:
\emph{"You build on dead ground," she says. "I can ward it---for a price. The firstborn child of your line, or a memory of the empire, or a promise to be named later." She smiles. "The soup is free. The price never is."}

\textbf{Options:}
\begin{itemize}
    \item \textbf{Firstborn.} Gain \textbf{Tarling Favor +2} (the ward works), but the witch marks the line. Future generations will be haunted.
    \item \textbf{Memory.} A PC sacrifices a cherished recollection (lose 1 XP, or a Bond). Gain the ward.
    \item \textbf{Promise.} Accept \textbf{Obligation [6]} to the witch. She will collect during the fourth sack.
\end{itemize}

\subsection*{The Tarling Mockery}
The local Tarling lord arrives for a "progress." He demands the Valvanos perform a play reenacting the surrender of the Tarlings to the empire.

Read:
\emph{"You will play the emperor," he says, handing you a paper crown. "I will play my own grandfather. We will kneel to you. And then you will kneel to me. The circle completes."}

If the party refuses, \textbf{Tarling Favor --4}. If they accept, \textbf{Tarling Favor +2} but \textbf{Valvano Loyalty --2}. A PC may deflect with a \textbf{Presence + Performance} (\textbf{DV 4}) to turn the mockery into a farce that humiliates the Tarling instead (requires \textbf{Valvano Loyalty 3+}).

\subsection*{The Ykrul Scouts}
A small Ykrul warband appears at the edge of the fen. They are not hostile.

Read:
\emph{"We remember the braid," their leader says. "We also remember a debt unpaid. The fourth sack is coming. Not today. Not tomorrow. But your grandchildren will see it."}

They leave a single horsehair braid tied to a stake before vanishing into the mist.

\section*{The Naming}
At the spring moot, the Tarling duke demands the cadet branch announce its new name.

Read:
\emph{The duke leans forward on his throne. "Well? Are you Valvano still? Or something else?" Aqyl lays a hand on your shoulder. "The ledger is open," he murmurs. "Write."}

\textbf{The Choice:}
\begin{itemize}
    \item \textbf{Valvano.} \textit{"We are the heirs of the Laurel. We will not kneel."} \textbf{Valvano Loyalty} must be 5+. Gain no land. The party must flee east or join the Ykrul.
    \item \textbf{Fenwood.} \textit{"We are of the fens and woods. The old name is ash."} \textbf{Tarling Favor} must be 5+. Gain the county, but the Everflame coal must be extinguished in public.
    \item \textbf{Fenwood-Valvano.} \textit{"We are Fenwood of Valvano."} Requires a compromise: \textbf{Valvano Loyalty} and \textbf{Tarling Favor} each at 3+. Gain the county, keep the coal, but the usurper will hear of it (\textbf{Obligation 2} to the usurper's agents).
\end{itemize}

\section*{Act IV Resolution}
Record the name. Aqyl speaks his final words:

\emph{"The fourth legacy is not the name. It is the choice. I have watched houses crumble for a thousand years. The ones that survive learn to bury their dead without burying themselves."}

He walks into the mist. No one sees him again.

\clearpage

\chapter*{Epilogue: The Braid and the Bell}
\addcontentsline{toc}{chapter}{Epilogue: The Braid and the Bell}

\lettrine{G}{enerations} later. The Fenwood keep stands. The Tarlings are gone---deposed by the Everbloods, with Fenwood swords. A child finds an old braid in the cellar---Ykrul horsehair woven with a Valvano seal. An old ledger records: \textit{"They left us a son. We sent him away. The Fenwoods do not speak of him."}

Outside, the Ykrul return for the fourth sack. But the Everbloods are now kings. And the Fenwoods are their trusted chancellor.

\vspace{2em}
\noindent\textit{The road does not forgive a debt. It only reschedules it.}

\vspace{2em}
\noindent\textit{--- From the Tarlington archives, copied in the year 876 AR, under the lamp, bread and salt. The ink is ash and rain.}

\clearpage

\chapter*{Appendix: Followers & Assets}
\addcontentsline{toc}{chapter}{Appendix: Followers & Assets}

\section*{The Unnamed Child (Half-Ykrul)}
If the party took the child from the Grey Ash camp, he becomes a follower.

\textbf{Cap: 2} \quad \textbf{Upkeep: 1} \quad \textbf{Asset: The Braid Token}
\begin{itemize}
    \item \textbf{Independent Action (once per session):} \textit{"He knows the steppe."} --- Gain \textbf{+2 dice} on any Survival roll involving tracking or foraging.
    \item \textbf{Harm:} 2 \quad \textbf{Exposure:} 2
\end{itemize}

\section*{The Debt Ledger (Heirloom)}
\textbf{Standard Asset (8 XP)} --- \textit{"Every oath recorded."}
\begin{itemize}
    \item \textbf{Free Use (once per session):} Ask the GM one question about an outstanding debt owed to the Valvano line. Gain a concrete clue.
    \item \textbf{Boon Activation (1 Boon):} Declare a minor debt that a local NPC owes. They provide a small service (food, shelter, information) without question.
\end{itemize}

\section*{The Everblood Promise}
If the party accepted the Linnic captain's offer, they gain a \textbf{String} on the future Everblood dynasty. Once per arc, they may cash this String to summon a small company of Linnic mercenaries (Cap 3) for one scene. After use, the String is consumed.

\clearpage

\chapter*{Quick Reference: Clocks}
\addcontentsline{toc}{chapter}{Quick Reference: Clocks}

\begin{tabular}{|l|l|l|l|}
\hline
Clock & Size & Advances When & Filled Effect \\
\hline
City Collapse & 6 & Failed recovery, GM SB spend & Gate seals; sewer escape required \\
\hline
Valvano Loyalty & 6 & Imperial rites, refusing humiliation & Cannot abandon the name; locked into Valvano or Fenwood-Valvano \\
\hline
Survival & 8 & Lost supplies, failed foraging, harsh choices & An NPC dies of starvation \\
\hline
Tarling Favor & 6 & Accepting humiliation, converting, gifting heirlooms & Granted the county (Fenwood ending) \\
\hline
Patience of the Gate & 4 & Failed trials, refused terms & Expelled from Vhasia; must seek Ykrul aid \\
\hline
Winter Provisions & 8 & Failed weekly survival rolls & Starvation claims a follower \\
\hline
\end{tabular}

\clearpage

\chapter*{Quick Reference: Heirlooms & Tokens}
\addcontentsline{toc}{chapter}{Quick Reference: Heirlooms & Tokens}

\begin{tabular}{|l|l|}
\hline
Item & Effect \\
\hline
Laurel Seal & +1 die to social rolls when presented \\
\hline
Bell Shard & Once per scene, make all allied rolls Dominant for one exchange \\
\hline
Debt Ledger & Asset; grants clues and minor service \\
\hline
Sacred Coal & Bonus with Everflame clergy; heretical to Temple of Light \\
\hline
The Braid Token & Ykrul favor; used in future adventures \\
\hline
Everblood Promise & String; summons Linnic mercenaries once \\
\hline
\end{tabular}

\clearpage
\chapter*{Colophon}
\addcontentsline{toc}{chapter}{Colophon}

\lettrine{T}{his} book was written in a fen-side hut, on paper made from rags and moss. The ink is lamp-black cut with the ash of a burned Tarling decree. The binding is horsehair.

The road does not forgive. The road does not forget. The road reschedules.

\textit{--- The Grey Wanderer, at the edge of Ash-Fenn, the year the braid was found.}

\vspace{1em}
\noindent The ninth cup remains un-poured. The ninth name remains unspoken.

\end{document}
