\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{fancyhdr}
\usepackage{multicol}
\usepackage{titling}
\usepackage{setspace}
\usepackage{hyperref}

≥ometry{margin=1in}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\titleformat{\section}{łarge\bfseries\uppercase}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}
\titleformat{⫕section}{ℹtshape}{}{0em}{}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Iron and Blood: The Ykrul}
\fancyhead[R]{þepage}
\fancyfoot[C]{Fate's Edge Module}

≠wcommand{\playmode}[1]{\subsection*{#1}}
≠wcommand{\culture}[1]{\subsection*{#1}}
≠wcommand{\gift}[2]{\textbf{#1:} #2}
≠wcommand{\timer}[2]{\textbf{#1 [#2]}}

\title{Iron and Blood: The Ykrul \\
łarge A Fate's Edge Module for Open−Plain Warfare, Logistics, and Diplomacy}
\author{}
\date{Version 0.3}

\begin{document}
\maketitle

\begin{quote}\footnotesize
\emph{On terms:} some Ecktorian broadsheets and street cant use ``orc'' as a slur for Ykrul communities. In play, treat it as prejudice with consequences. The peoples name themselves \textbf{Ykrul}, with many local kin−names besides.
\end{quote}

\vspace{−0.5em}
\section*{Design Goals}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Reputation and reality:} Ykrul bank on a fierce reputation on the open plain−−−and then win by routes, supply, and parleys that make blades unnecessary.
  ℹtem \textbf{Plural lives:} herders, pilots, weavers, judges, scouts, factors, captains; warfare is one tool among many.
  ℹtem \textbf{Kon'reh in the marrow:} the board's geometry informs roads, exits, concessions, and seasonal rights.
  ℹtem \textbf{Core−first:} Position/DV/Effect, timers, and Strings do the work; no new math.
\end{itemize}

\section*{Quickstart (2 minutes)}
\begin{enumerate}
  ℹtem Choose your lens: \textbf{Road−Clan PCs} • \textbf{Mixed Company} • \textbf{Outsiders Among Ykrul}
  ℹtem Pick a home ground: \textbf{Meadow Commons} • \textbf{Mountain Holds} • \textbf{Salt Coasts} • \textbf{Eastern Steppes}
  ℹtem Mark \textbf{Repute [6]}, \textbf{Standing [6]}, \textbf{Debts [4]} (opt. \textbf{Marks [6]} for toil/weather)
  ℹtem Choose \textbf{2 Kin Gifts} and \textbf{1 Bond} (trade, craft, law, or defense)
  ℹtem Start with \textbf{1 String} (e.g., windbreak right) and \textbf{1 Obligation} (e.g., ford dues)
  ℹtem Map 4−−6 \textbf{nodes} and light local timers: \timer{Market Moot}{4}, \timer{Winter Drive}{6}, \timer{Border Ride}{6} (swap in \timer{Blood−Feud}{4} only if fiction calls for it)
  ℹtem Open with a situation: \textbf{Passage Negotiation} • \textbf{Winter Drive} • \textbf{Harbor Dues} • \textbf{Kon'reh Arbitration}
\end{enumerate}

\section{Ykrul −−− The People of the Violet Steppe}

\subsection*{Kon'reh as sacred geometry, memory of the Wake, and a world seen in lines, weights, and exits}

\begin{quote}
``Through the Wake, Across the Meadow, Into the Stone.''
\end{quote}

\medskipℎrule\medskip

\subsection{I. The Great Migration (Wake−Law)}

Ykrul memory begins with flight: homesteads abandoned under a wrong sky, flotillas stitched from last ropes, and an ocean learned by star and sound. They call that crossing the \textbf{Great Wake}. Every clan keeps a coil of salt−stiff rope−−−the \textbf{Salt Line}−−−uncoiled before serious speech to place all present ``under Wake−law'': plain words, straight debts, no riddles.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{First Meadow.} Landfall was a long, violet grassland. Its color−−−ironweed, crocus, mallow−−−named their world: the \emph{Violet Steppe}.
  ℹtem \textbf{Wake Names.} Many wear a second name granted by the sea (\emph{Wave−Left}, \emph{Stern−Watcher}), which can be invoked for one unlooked−for courtesy ``as kin on the crossing.''
  ℹtem \textbf{Rites that stuck.} A handful of salt in a new road's rut; a black river stone atop an elder's cairn; a song that lists exits before listing enemies.
\end{itemize}

\emph{How they think:} the Wake taught them that survival comes from reading \textbf{flow} (what moves), \textbf{weight} (what holds), and \textbf{exit} (what opens). Kon'reh will codify this−−−but the Wake is the first proof.

\medskipℎrule\medskip

\subsection{II. Lands After Landfall}

⫕section*{The Violet Steppe (Meadow)}
Wind, hooves, and shallow lakes. Herd−bands form \emph{ring−camps} (tents inward, cookfires outward); visitors call at the outer fire and announce themselves. Major decisions take place ``by \textbf{Bowl and Board}'': the \emph{Bowl} to test plain fairness, the \emph{Board} to model routes and concessions.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Strings of the Meadow:} rain−marker rod; herd−tally braid; windbreak right; salt share.
  ℹtem \textbf{Persistent worries:} black gnats; rivers that change their beds overnight; border rides that test how far a ring will bend.
\end{itemize}

⫕section*{The Eastern Steppes}
Vaster, colder, older. \emph{Sky cairns} mark wind and weather; riders replace a fallen stone and leave an ember as courtesy to strangers and spirits. Ward−storms roll in sheets; Kon'reh analogies (flow/weight/exit) double as actual storm−reading.

⫕section*{Mountain Holds}
Knife ridges and pass−stones. Mountain Ykrul are infamous even among Aeler captains: hard meat, harder hospitality. They swear \textbf{Stone−Oaths} at marked passes and genuinely believe a broken promise wakes the ridge (and avalanches have a long memory). Climbing gear is braided rope and iron wrapped in cloth; iron bares teeth only in war.

⫕section*{Salt Coasts}
Sea Ykrul cut sleet−thin wake−cutters and braid hullcraft with Linnic tricks. Harbor politics oscillate between \textbf{Blue Moots} (truce and tariff) and dawn raids. They swear by the \textbf{Three Knots} (helm, mast, rudder) and settle maritime disputes at a beach−board chalked into wet sand.

\medskipℎrule\medskip

\subsection{III. Kon'reh: Sacred Geometry Lived}

Kon'reh is both a table game and a theology of place: the \emph{Meadow} (pressure and path), the \emph{River} (flow and change), the \emph{Stone} (weight and witness). The \emph{Concordance}−−−commentary cycles on Kon'reh and cosmology−−−treats roads, rivers, and ridges as the language of intention. Ykrul do not separate ``strategy'' from ``ethic''−−−to place well is to behave well: no route you can't defend, no exit you build only to close behind weaker feet.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Bowl & Board.} Public quarrels are fixed first by fairness (Bowl), then by geometry (Board). A concession on the Board becomes a season−long right at a ford, gate, or pasture.
  ℹtem \textbf{Schools (as temperaments).}
    \begin{itemize}
      ℹtem \emph{Meadow School:} control the middle by promising exits you can afford.
      ℹtem \emph{Mountain School:} weld lanes until only one honest path remains.
      ℹtem \emph{Wake School:} trade tempo like wind; arrive before you seem to move.
      ℹtem \emph{Aeler−influenced ledgers:} price the route−−−make bad paths expensive rather than forbidden.
    \end{itemize}
  ℹtem \textbf{Devotion without temples.} Shrines are boards carved in stone, lanes picked in cairns, and field−camps marked by ring and windbreak. The sacred is what holds under weight and gives under wisdom.
\end{itemize}

It's not their only game. Ykrul play \emph{river dice} (eddy−based gambling), \emph{knot puzzles} (one−breath releases), \emph{horse−circles} (precision riding), and \emph{salt−bones} (story−sticks cast in patterns). But Kon'reh is how elders teach children to think about promises, borders, and the dignity of ground.

\medskipℎrule\medskip

\subsection{IV. Law, Diplomacy, War−−−Seen in Lines}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Guest Right.} A guest cup at the outer fire; theft ``under smoke'' brands your tent marks gray for a season.
  ℹtem \textbf{Two Ledgers.} Said and meant are recorded together; offering both averts face−losing traps.
  ℹtem \textbf{Blood−Price.} Paid in animals, salt, or length of rope; a band that refuses enters \emph{Red Weather} (others treat them as if a storm hangs over their camp).
  ℹtem \textbf{Silence Furlong (with the Lethai−al).} A speechless border strip: no grazing, no felling, no names. Cross in silence, then speak once. Over time this produced mutual isolation and a brittle respect; when the strip is honored, councils run warm−−−when not, gray−fletched messengers appear at dusk.
  ℹtem \textbf{Pass & Harbor Doctrine.} In mountains, hold what you must and never promise what stone will not bear. At sea, get there first or sing the storm together; a shared song is worth more than a sharp keel.
\end{itemize}

\medskipℎrule\medskip

\subsection{V. People & Institutions}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Meadow Judges.} A traveling trio who carry a bowl, a board, and a braid; their decision stands one season and one road.
  ℹtem \textbf{Kon'reh Masters.} Teachers who arbitrate by geometry; respected even by rivals. They rarely smile, often bleed for other people's mistakes, and will play anyone who brings a decent board and a decent reason.
  ℹtem \textbf{Stone−Sons & Rope−Daughters.} Mountain societies that test initiates on night crossings. Their braids anchor Aeler engines and their oaths make Aeler officers polite.
  ℹtem \textbf{Wake−Wrought.} Sea families that list their hulls like saints; name the rivets of a stolen boat and you can claim it back under Wake−law.
\end{itemize}

\medskipℎrule\medskip

\subsection{VI. Aesthetics & Speech}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Colors:} violet (meadow), gray (stone), blue−white (wake).
  ℹtem \textbf{Shapes:} rings, crosses, and stepped lines; carpets and tattoos look like maps.
  ℹtem \textbf{Proverbs:}
    \begin{itemize}
      ℹtem Leave two exits; take one.
      ℹtem If the river bends, your story should too.
      ℹtem A promise that cannot bear weight is a trap, not a vow.
      ℹtem Never place what you will not defend.
    \end{itemize}
  ℹtem \textbf{Names:} given + road + deed: \emph{Sera of the Third Road, Who Turned the Herd}. Hand−signs say policy faster than words: palms down = camp law; fingers split = board; knuckles together = witness shut.
\end{itemize}

\medskipℎrule\medskip

\subsection{VII. Rivals & Kindreds}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Lethai−al (wood elves).} Ancient skirmishes shaped strict courtesy. The forest never forgets; the grass never lies. Councils hover between poetry and border drills.
  ℹtem \textbf{Aeler (dwarves).} Mutual admiration where work meets weather. The Rope and the Ring−−−each holds where the other cannot.
  ℹtem \textbf{Linns (mist & iron).} Kinship of hull and hunger; borrowed sailcloth and storm−songs; first to the river mouth takes the season unless a Blue Moot is sworn.
\end{itemize}

\medskipℎrule\medskip

\subsection{VIII. Festivals & Omens}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Banner Day.} Spring match between bands: not about victory so much as who sets summer routes.
  ℹtem \textbf{Blue Weather Choir.} Storm season rite where ships and camps sing exit−lists in antiphon with thunder.
  ℹtem \textbf{Gray Hand Month.} When a band has failed blood−price; no mummers dance their fires.
  ℹtem \textbf{Omens that matter:} violets closing at midday; cairn shadows pointing backward; a wind that tastes of copper; horses all looking east at once.
\end{itemize}

\medskipℎrule\medskip

\subsection{IX. Adventure Seeds}

\begin{enumerate}[leftmargin=*]
  ℹtem \textbf{The Bowl at the Ford.} Meadow Judges convene; a toll war brews between Sea Ykrul and Linns. Your party may broker a concession that becomes a season−long ford right.
  ℹtem \textbf{Stone−Oath Winter.} A pass promise broke itself; avalanches argue like lawyers. Reset the oath before a rival Aeler captain ``fixes'' the ridge with iron.
  ℹtem \textbf{Silence Furlong, Broken.} Someone spoke and cut grazing under the roof−trees; restore trust with a geometry both sides can live in.
  ℹtem \textbf{The Stolen Board.} A Kon'reh Master's stones−−−heirlooms from First Meadow−−−are missing on the eve of a marriage pact. Recover them or design a substitute set both clans will accept.
  ℹtem \textbf{Blue Weather Choir.} Ward−storms roll from the Eastern Steppes; sing with Sea Ykrul and Linn captains−−−or watch the harbor tear itself in half.
\end{enumerate}

\medskipℎrule\medskip

\subsection{X. Portraying the Ykrul}

\begin{itemize}[leftmargin=*]
  ℹtem Let geography talk first: route, exit, weight, witness.
  ℹtem Treat Kon'reh as grammar, not hobby−−−how they make an argument about roads, borders, and mercy.
  ℹtem Use Bowl then Board to structure any dispute: fairness $→$ geometry $→$ season−long concessions.
  ℹtem Keep the Wake close: salt and rope as ritual memory; plain words when stakes are high.
  ℹtem With rivals, play courtesy like a blade−−−sharp, exact, and meant to leave everyone standing if they choose.
\end{itemize}

\emph{For the Ykrul, Kon'reh is not simply a pastime. It is the proof that the world has form−−−and that people owe each other exits.}

\section{The Ykrul Ways (four grounds)}
\culture{Meadow Commons}
\textbf{Where:} violet grass, ring−camps, shallow lakes. 
\textbf{Known for:} fast musters, moving markets, distance diplomacy. 
\textbf{Kin Gifts (pick 2):} Herd−Mastery; Weather−Reading; Caravan Craft; Route−Planning (Kon'reh framing). 
\textbf{Strings:} rain−marker rod; herd−tally braid; windbreak right. 
\textbf{Timers:} \timer{Black Gnat Season}{6}; \timer{Border Ride}{6}.

\culture{Mountain Holds}
\textbf{Where:} pass−stones, winter bowls, knife ridges. 
\textbf{Known for:} holding lines, winter stores, signal sense. 
\textbf{Kin Gifts:} Stone−Sense; Counterweight Engineering; Rope−Craft; Avalanche Reading. 
\textbf{Strings:} pass−stone mark; winter fodder charter. 
\textbf{Timers:} \timer{Winter Drive}{6}; \timer{Ridge Negotiation}{4}.

\culture{Salt Coasts}
\textbf{Where:} rocky inlets, island runs, river mouths. 
\textbf{Known for:} pilots, moots, storm windows. 
\textbf{Kin Gifts:} Storm−Seamanship; Harbor Dues; Shoal Mapping; Blue Moot Etiquette. 
\textbf{Strings:} harbor priority; wake−rivet list. 
\textbf{Timers:} \timer{Harbor Blood}{4}; \timer{Blue Weather}{6}.

\culture{Eastern Steppes}
\textbf{Where:} sky cairns, ward−storm belts, long legs. 
\textbf{Known for:} exit−finding, omen−reading, cross−cultural guides. 
\textbf{Kin Gifts:} Cairn−Talking; Ward−Storm Guidance; Long−Leg Logistics; Silence Furlong Etiquette. 
\textbf{Strings:} cairn ember; guide−right. 
\textbf{Timers:} \timer{Ward−Storm Track}{4}; \timer{Cairn Misread}{4}.

\section{Repute, Standing, Debts}
\textbf{Repute [6]} −−− kept bargains, delivered goods, exits honored. 
0−−2: distrusted (Position −−1 in negotiation) • 3−−4: reliable • 5−−6: sought−after (Position +1, tempting target).

\textbf{Standing [6]} −−− weight in councils and moots. 
0−−2: marginal • 3−−4: contributor • 5−−6: chair/warden (responsibility timers light).

\textbf{Debts [4]} −−− dues, oaths, promises. 
At full: dues come due (court, kin, or market).

\section{How Ykrul Win (beyond the blade)}
\subsection*{Banked Fear (Reputation Economy)}
On entering a venue where your fierce name precedes you, mark \textbf{Banked Fear (1)}.\\
\emph{Spend 1:} force \textbf{Parley First} (one roll of talk before a fight can start) \emph{or} shift one enemy action to \textbf{Controlled} (they flinch at your name).\\
If you bluff and then fail to deliver, erase all Banked Fear until you win publicly.

\subsection*{Logistics Edge (Strings \textrightarrow\ DV)}
Convert any \textbf{cache String} (fodder lot, water right, hidden wharf) into \textbf{DV −−1} on a relevant leg or score. If used to starve an innocent district, flip it: DV −−1 becomes a public \textbf{Debts +1}.

\subsection*{Kon'reh Arbitration (Geometry of Mercy)}
Model roads, exits, and lanes in talk. On success, create a \textbf{Seasonal Concession String} at a ford/harbor/pass that either side may call once/season without offense.

\section{The Vilikari Compact (four autumns)}
\subsection*{A ledger written in other hands}
The Ykrul and Vilikari have fought, married, traded, and cheated together across three coasts. The Syndicate likes coin that moves; the Ykrul like routes that hold. Between them grew the \textbf{Compact}: when Ykrul choose \emph{where} and \emph{when}, Vilikari decide \emph{how hard}.

⫕section*{The Four Sacks of Ecktoria}
Older Ykrul speak of \textbf{four autumns} when Ecktoria's pride buckled. The banners were Ykrul. The bulk of the street−fighting were Vilikari crews−−−bought dear, pointed like a storm, paid on the steps. Quartermasters with Kon'reh boards widened alleys into lanes, closed gates with carts, and fed the fists who did the breaking. The lesson stuck: \emph{win the routes; hire the hands}. It left Ecktoria wary, the Vilikari rich, and the Ykrul rumor larger than their numbers.

⫕section*{Table Levers (Compact)}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Hire the Hands, Own the Routes:} pay a Vilikari cohort with coin or cut; gain a \textbf{Fists [3]} asset and \textbf{Route Priority (1)} at one gate/ward for the score.
  ℹtem \textbf{Two Ledgers, One Purse:} keep a clean ledger (said/meant) while paying dirty; on a strong hit, convert one \textbf{Crew Heat} into \textbf{Debts +1 (Vilikari)} instead of City Heat.
  ℹtem \textbf{Autumn Memory:} in Ecktoria−only scenes, your name grants \textbf{Banked Fear (1)} even if unknown individually.
\end{itemize}

\section{Strings, Secrets, Nodes}
\textbf{Strings:} windbreak right; pass−stone mark; harbor priority; board−stone concession; wake−rivet list; cairn ember.\\
\textbf{Secrets:} hidden ford; old toll clause; moot−fixing scandal; Syndicate cut; winter cache.\\
\textbf{Nodes (pick 4−−6):} Violet Meadow • Market Borough • Border Ford • Stone Hold • Sky Cairn • Coastal Harbor • Island Wharf • Rope−Bridge • Ring−Camp.

\section{Procedures & Ladders (examples)}
\textbf{Passage Rights} DV 2 posted • 3 surge • 4 contested • 5 embargo.\\
\textbf{Harbor Dues} DV 2 public rate • 3 guild day • 4 faction squeeze • 5 blockade.\\
\textbf{Compact Hire} DV 2 desperate crews • 3 standard • 4 disciplined • 5 elite union.\\
\textbf{Kon'reh Talk} success = \emph{Seasonal Concession String}; miss = \emph{Market Moot [4]} advances.

\section{Regional Notes (quiet variety)}
\subsection*{Mid Ahkaz}
A city that learned to breathe in seasons: fairs that walk, courts that listen before they price, and neighborhoods where Ykrul calendars hang beside local saint−days.

\subsection*{Ecktoria}
Stone streets remember hoofbeats. Some doors open, some spit. Hirelings keep their eyes on the purse and the alley mouth. The Syndicate keeps longer books than it admits.

\subsection*{Vhasian & Viterran Cities}
River towns price the tide; hill markets price the view. A banner can get you an audience; a good ledger gets you paid.

\section{Adventure Frames}
\subsection*{The Fifth Autumn (Social Heist)}
A Vilikari bloc wants to replay an old triumph in Ecktoria. You intend to \emph{not} sack anything: win a tariff by making the threat obvious and the routes yours. Bank fear, hire hands, feed them well, and leave with a contract.

\subsection*{Moot of Blue Weather (Diplomacy)}
Storm season and a choke at the river mouth. Call a Blue Moot, keep Linn captains in the room, convince Syndicate brokers the rate holds, and make the tide window safe for all without giving away the lane.

\subsection*{Silence Furlong, Mended (Frontier)}
A speechless strip on the border frayed. Walk it clean, pay old dues, set a seasonal concession stone both sides can live with, and escort a caravan through without drawing.

\section{Example of Play}
\textbf{Setup:} Meadow negotiators (Caravan Craft, Route−Planning) and coastal pilots (Harbor Dues) work an Ecktorian tariff crisis. A Vilikari bloc offers fists if the purse is right.

\textbf{Play:} Banked Fear earns \emph{Parley First}. Kon'reh Talk frames a seasonal lane and exit. Compact Hire brings disciplined crews who never need to swing−−−their presence is enough once the food and coin are sure. Logistics Edge burns a cache String for DV −−1 on the night's movements. By dawn, the city posts a rate. No sack, no siege; a contract and a story.

\textbf{Outcome:} \emph{Market Moot [4]} clears; \emph{Seasonal Concession String: harbor priority} created; \emph{Debts +1 (Vilikari)} noted for the purse that made it all graceful.

\section{Aveh −−− The Rider Behind the Storm}

\subsection*{Non−binary patron of freedom, edges, and the unclaimed road}

\begin{quote}
``No road owns me. No name cages you. Ride.''
\end{quote}

\medskipℎrule\medskip

\subsection{I. Names, Faces, Signs}

Aveh rides where borders fray: along stormfronts, river booms, city rooftops, ridgebacks, and the hush between a question and an answer. Neither man nor woman (and not ``both'' so much as \emph{other}), Aveh is called when people refuse a cage or cannot fit the one built for them.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Epithets:} The Unreined; Way−Mercy; Break−Reins; The Door in the Wind; The Rider Without Shadow.
  ℹtem \textbf{Signs:} an inverted bridle; a broken bit−ring; a wind−bell singing with no air; a spiral cloud with an empty eye; footprints that begin mid−path.
  ℹtem \textbf{Shrines:} roof−altars with a single unlit lamp; bridle rings on cairns; chalk spirals at ferry posts; wind−bells hung under eaves.
\end{itemize}

\medskipℎrule\medskip

\subsection{II. The Road's Mercy (Doctrine)}

Aveh's teaching is simple and sharp, like rain on hot stone.

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{No coercion.} Do not bind a will without its consent; do not be bound without yours.
  ℹtem \textbf{Leave exits.} Every plan keeps a way out for the least of you.
  ℹtem \textbf{Break bad bindings.} Vows made under harm or lie are weather to be ridden through, not walls to revere.
  ℹtem \textbf{Keep witness gentle.} Names are gifts; use the ones people choose.
\end{itemize}

Among Ykrul, this reads as the conscience inside sacred geometry: a route is only honest if those who must walk it can refuse.

\medskipℎrule\medskip

\subsection{III. Shrines, Rites, and Everyday Practice}

\textbf{Freehouses} (door−bells, unlatched latches) are Aveh's favored shrines in towns and on roads. Custom says: \emph{no questions first, names after consent, doors unlatched unless danger presses.}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{The Open Name.} Speak the name you choose; listeners answer, ``Ridden and riding,'' acknowledging your authorship of self.
  ℹtem \textbf{The Unknotted Door.} Loosen one knot a day: a stuck latch, a tangled cord, a bureaucratic snare for a stranger.
  ℹtem \textbf{Storm Cup.} Keep a spare cup by the door; when the unlooked−for arrives, pour first, ask later.
  ℹtem \textbf{Rite of Unnaming.} Ash on the brow, rainwater over the tongue: ``Unname the harm, keep the memory.'' Used to shed imposed labels while preserving history.
  ℹtem \textbf{Break−Reins Feast.} At first storm of the season, bits are hung on a thorn or eave; those who cast off false ties pass beneath wind−bells and are applauded without questions.
\end{itemize}

\medskipℎrule\medskip

\subsection{IV. Adherents & Households}

\textbf{Nameless Riders} are wandering mystics of Aveh: weathered cloaks, a bell, and writs that petition courts or clans to release people from unjust bonds. Some practice \emph{total unnaming} for a season; wiser ones keep one thread (a color, a song) to anchor self.

\textbf{Households of the Open Door} are ordinary faithful: shelter−keepers, caravan hands who smuggle people rather than goods, clerks who ``lose'' bad papers, pilots who wave through the right boat on the wrong day. Chosen families flourish here.

\medskipℎrule\medskip

\subsection{V. Law, Courtesy, and Crossroads}

Aveh does not despise law; Aveh despises \emph{coerced} law. In Ykrul country, \emph{Bowl before Board} becomes \emph{Open Name before Bowl}: the courtesy of naming precedes fairness, which precedes geometry. Harbor captains hang wind−bells during \emph{Blue Moots} to signal that refugees and misfits pass first. In courts fond of traps, Aveh's folk keep two ledgers (said/meant) and make both explicit.

\medskipℎrule\medskip

\subsection{VI. Omens & Visitations}

A door that will not latch on a still night; a sudden scent of rain indoors; a bridle ring in a gutter; a lantern that refuses to light until a question is asked kindly. Aveh appears as a rider without reins, a voice like wind through reeds−−−precise, amused, never cruel.

\medskipℎrule\medskip

\subsection{VII. Tokens & Strings (in play)}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Wind−Bell (sanctuary).} Signals a place where the unlatched custom holds; once/scene, improve Position when seeking safe audience or exit.
  ℹtem \textbf{Bridle Ring (oath−breaker).} Voids one coercive clause in a negotiation; converts it into a public concession timer.
  ℹtem \textbf{Name−Shard.} Establishes chosen−name etiquette in a scene; disrespect carries social harm.
  ℹtem \textbf{Storm−Scrip.} A small bag of cut cords; cash to soften a blockade or bypass petty gatekeeping without blood or bribe.
\end{itemize}

\emph{Optional table norm:} When a PC protects someone's autonomy at a cost, grant \textbf{Storm's Favor (1)}; spend to shift Position/DV in escapes, parleys, and boundary crossings that harm no bystander.

\medskipℎrule\medskip

\subsection{VIII. Cross−Cultural Notes}

\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Ykrul.} Aveh is the exit written into the board: never place what you will not defend, never close the last way out.
  ℹtem \textbf{Aelinnel.} Two ledgers (said/meant) become sacrament; copper over iron in fraught negotiations.
  ℹtem \textbf{Aelaerem.} Guest−right tilts toward the stray: bread & salt first, genealogy later.
  ℹtem \textbf{Linns.} Mast−bells ring in storm season; ropes are cut before hulls become prisons.
\end{itemize}

\medskipℎrule\medskip

\subsection{IX. Adventure Seeds}

\begin{enumerate}[leftmargin=*]
  ℹtem \textbf{The Door That Wouldn't Latch.} A latch refuses to hold; behind it, papers forged under duress. Fix the bind, not the door.
  ℹtem \textbf{Break−Reins at First Storm.} A contract hides a coercive clause; ride the festival's chaos to unmake it without blood.
  ℹtem \textbf{Unbinding Letters.} Three border posts; three different lawful traps; one caravan carrying freedom writs.
  ℹtem \textbf{The Nameless Rider.} A mystic erased so much that their allies forgot the revolution they began. Find the anchor and give the name back.
  ℹtem \textbf{Blue Moot Without Toll.} Harbor captains demand papers from refugees; persuade them to declare a toll−free moot for one night−−−and hold the line when profiteers push back.
\end{enumerate}

\medskipℎrule\medskip

\subsection{X. Portraying Aveh}

Lead with courtesy that opens space. Make names holy and doors unlatched unless danger presses. Let storms change the scene \emph{toward mercy}. When in doubt, ask: \emph{Where is the exit, and who needs it most?}

\quotebox{``If the empire wants a receipt, we bring one. If the steppe wants an exit, we leave two. If the banners want pay, we count it twice.''}

The Vilikari are natives of Ecktoria's river valleys and the Violet Steppe fringe, living where forts end and pasture begins. To imperial clerks they're ``barbarians'' (an exonym with teeth); to caravaners and captains they're the people who can make a bad road pay by morning. Their reputation−−−hustlers, fixers, smugglers−−−isn't a destiny so much as a practiced craft: reading law and landscape at the same time.

They sit at three tables at once:
\begin{itemize}
ℹtem with Ykrul judges over Bowl and Board,
ℹtem with Ecktorian customs under lamps and ledgers,
ℹtem with Black Banner mercenaries in counting−tents that smell of oil and pay−chest wax.
\end{itemize}

When the Ykrul ``sacked Ecktoria'' four times, it was often Vilikari cohorts who opened gates, fed horses, and brokered pay−−−route victories more than field ones.

\section{Origins & Shape}

\subsection{After the Wake, Before the Walls}

When the first Meadow bands pushed west, some families settled where ferry rights and pasture dues interlocked. Those families−−−later called Vilikari−−−built ring−forts on hill spurs, river quarters under bridges, and road−camps that could fold in an hour. They learned imperial paper because it moved grain; they learned Kon'reh because it moved people.

\subsection{Three Houses, One People}

\begin{description}[leftmargin=*, labelwidth=2cm]
ℹtem[Reed−House (rivers)] barge teams, lock−keepers, night ferries; speak in tariffs and tide tables.
ℹtem[Stone−House (hills)] ring−forts, dry−stone craft, winter markets; speak in oaths and neighbors.
ℹtem[Road−House (camps)] caravans, remount strings, border paths; speak in exits and times.
\end{description}

Expect intermarriage and seasonal migration between the three; expect cousins in every garrison town from Mid Ahkaz to the Ecktorian road hubs.

\section{Law, Hustle, and How They Win}

\subsection{Reed−Law (bend, don't break)}

A ledger you can bend is worth more than a bridge you can't cross. Vilikari law centers on procedural mercy: show the rule, then the exception, then the witness who makes it stick.

\subsection{Night Roads}

The Vilikari art of moving goods or people through legitimate gaps−−−after bell, before dawn, with a paper that's almost right. It's not crime by default; it's a relief valve for systems that jam.

\subsection{Kon'reh as streetcraft}

Ykrul sacred geometry−−−flow/weight/exit−−−is applied to kiosks, toll lanes, and patrol routes. The Vilikari don't outfight a checkpoint; they re−price the route until the gate opens for someone.

\subsection{Black Banner Accounts}

Their long partnership with the Black Banner (a mercenary franchise) taught them payroll discipline and camp hygiene. In many campaigns the Banner provides muscle, the Vilikari provide supply, scouts, and brokers, and the empire provides the contracts−−−until the price changes.

\section{Etiquette (Mechanics−ready)}

Use these as Position/DV levers at the table; no new math.

\mechanic{Lamp−Lit Paper}{
Present any writ under a lit lamp and read it aloud−−−including the margins.\\
\textbf{Effect:} Position +1 on bureaucratic scenes; on a miss, start Imperial Audit [2].
}

\mechanic{Two Weights, One Truth}{
Lay out public and private prices together (``said/meant'').\\
\textbf{Effect:} DV −1 to close a deal; if you hide the second ledger, tick Night Road Heat [1].
}

\mechanic{Guest Cup on the Threshold}{
Pour for the guard before yourself.\\
\textbf{Effect:} First social SB against you can be converted to Heat +1 instead of a complication.
}

\mechanic{Banner Left Standing}{
In any parley with mercenaries, keep a spear or flag upright by the entrance.\\
\textbf{Effect:} Once/scene you may turn a melee into a negotiation without changing venues.
}

\section{Strings (Receipts you can spend)}

\begin{description}[leftmargin=*]
ℹtem[Reed Chit] A stamped lock−pass; once per session DV −1 on crossings or inspections.
ℹtem[Toll Braid] A woven cord showing last season's dues; Position +1 at toll houses.
ℹtem[Black Banner Voucher] IOU from a captain; cash to call 1 squad for 1 scene−−−or cancel one squad chasing you.
ℹtem[Guild Widow's Trust] A fence or factor will hold goods no questions asked; mark City Heat +1 on use.
ℹtem[Ford Right (Season)] A concession earned at Bowl & Board; treat a specific ford/bridge as friendly terrain.
\end{description}

\section{Timers & Fronts (Their world bites back)}

\begin{description}[leftmargin=*]
ℹtem[Night Road Heat {[}6{]}] Customs, syndicates, and jealous rivals. On fill: raids, curfews, embargoes.
ℹtem[Imperial Audit {[}4{]}] A clerk with a pen sharper than swords. On fill: seizures, fines, informant flips.
ℹtem[Banner Debt {[}4{]}] Missed payments or blown promises to mercenaries. On fill: a ``discipline visit.''
ℹtem[Ford War {[}6{]}] Competing rights over a crossing; each session the price to pass changes.
\end{description}

\section{Institutions & Everyday Faces}

\begin{description}[leftmargin=*]
ℹtem[Countkeepers] Vilikari auditors who can read both imperial tallies and meadow braids; hired by anyone who doesn't want to be cheated−−−including you.
ℹtem[Ferry Mothers] Night ferry syndics who control skiffs, bribes, and rumor flow; break their trust and you walk.
ℹtem[Gate Cantors] Professional ``readers−aloud'' who know which clause saves you; they carry their own lamps.
ℹtem[Camp Brokers] Middlefolk who place you with mercenary companies or get you out of them.
\end{description}

\subsection{Reputation as Tool}

``Hustler, smuggler, barbarian.'' The Vilikari let empire call them names because fear makes prices better. Anyone who's needed a quiet crossing learns quick who they really are.

\section{Relationships on the Edge}

\subsection{Ykrul}

Cousins by road. Vilikari serve as interpreters of Kon'reh to imperial courts and of paper to meadow judges.

\textbf{At the table:} Bowl → Board + Two Weights creates season−long rights everyone honors.

\subsection{Ecktoria}

Love−−hate. Vilikari run the empire's markets after dark and its fords at dawn. Clerks call for audits; generals ask for guides.

\textbf{At the table:} Spend a Reed Chit to turn a checkpoint into a transaction.

\subsection{Black Banner}

Familiar. The Banner trusts Vilikari quartermasters because feed and pay arrive on time. If Banner Debt fills, trust evaporates fast.

\subsection{Aeler & Lethai}

Aeler hire Countkeepers; Lethai trade seed−tithes for night ferry know−how. Vilikari often mediate keystone vs. hedge disputes for a fee.

\section{Running Vilikari PCs (Hats, not classes)}

\begin{description}[leftmargin=*]
ℹtem[Key−Bearer] Keeps the lamp, knows the clause. Once/scene, convert a hard ``No'' into a Project [4] (``Bring X papers, do Y repair'').
ℹtem[Road−Cutter] Finds legal gray lanes. Once/leg, reroute around a Front at the cost of Night Road Heat +1.
ℹtem[Bargainer] Turns reputation into leverage. On a strong hit in social scenes, take +1 Effect if you accept Notoriety [1].
ℹtem[Quartermaster] Makes scarcity obey. Once/score, downgrade a Supply consequence to a timer you can work.
\end{description}

\section{Scenes & Scores}

\begin{enumerate}
ℹtem \textbf{The Lamp and the Gate.} Customs closes a bridge after a scandal. Present lamp−lit papers, read the clause right, and offer a repair project instead of a bribe. If you fail, Imperial Audit +1 and the Ford War timer starts.

ℹtem \textbf{Pay the Banner.} A Black Banner captain is short three chests. You can smuggle coin, re−price the route with a market coup, or swap debt for a season of ford rights−−−but tick Banner Debt if you break your word.

ℹtem \textbf{Two Weights at Bowl & Board.} Ykrul judges and city clerks meet at a ford. You must show both ledgers and leave with a season−right everyone can live with. If you hide a number, Night Road Heat ticks and the ford becomes ``expensive.''

ℹtem \textbf{Night Ferry Gospel.} A famine town needs grain past curfew. Move the barges on the Night Roads without waking the Audit; spend Guild Widow's Trust to float the collateral.

ℹtem \textbf{The Gravel Choir.} A Stone−House ring−fort owes repair years after a skirmish. Broker labor between Aeler masons and Lethai wardens; convert a revenge plot into a repair ledger people will actually pay.
\end{enumerate}

\section{Look & Feel}

\begin{description}[leftmargin=*]
ℹtem[Dress] dark wool, bone toggles, lamp−glass charms; braids that show dues paid.
ℹtem[Sound] reed flutes at night, ledgers read aloud under lamps, the clunk of traveling weights.
ℹtem[Symbols] braided cords (paid/owed), ferry tokens, coin scales stamped with a starboard mark.
\end{description}

\subsection{Sayings}

\begin{itemize}
ℹtem ``Paper under lamp, promise under witness.''
ℹtem ``Two weights, one truth.''
ℹtem ``Leave two exits; we'll buy the cheaper.''
\end{itemize}

\s
\appendix

\section{The Ykrul Thirty−Six Sayings}

They measure war not in strategies, but in breaths left to their foe.

\begin{enumerate}
ℹtem \textbf{Let the River Freeze Before You Cross}
Do not rush from your Home Apex. A Blue that leaves too early is a calf separated from the herd. Let the opponent make the first shape; your bridge will be built upon their impatience.

ℹtem \textbf{Borrow a Corridor to Trap a King}
When the enemy carves a path toward your sanctum, let it deepen while you count his exits. Then seal the mouth behind him. Their ambition becomes their tomb.

ℹtem \textbf{Kill With a Borrowed Blade}
Your Reds are cheap. Their purpose is to be spent. Lure the enemy's Orange into capturing a Red, placing it where your Blue can hop over it and land clean on the empty square beyond to strike a greater prize.

ℹtem \textbf{Watch the Fire Burn Across the River}
When the enemy's Blue is Rooted after a Seed, do not panic. They have spent their breath to gain a piece. Use their immobility to post your lids and print your lanes. Let them burn their own tempo.

ℹtem \textbf{Loot a Burning House}
Strike only when the enemy is committed elsewhere. A Blue in the Cross is a house with one door. A Blue after its second special is a house on fire. Plunder it.

ℹtem \textbf{Hide Your Dagger Behind a Smile}
Telegraph a threat on one flank. When they shift their lattice to meet it, strike on the silent file they were forced to weaken.

ℹtem \textbf{Cross the Sea by Fooling the Sky}
Make a small, obvious move to hide a greater one. A Red slide of two squares can scream so loudly it deafens them to your Orange sliding three elsewhere.

ℹtem \textbf{Cull a Kid to Save the Herd}
A Green is a piece of tempo, not a piece of value. Let it be captured if it baits their Blue into a square from which it cannot escape your true trap.

ℹtem \textbf{Take the Opportunity While He Counts Another Man's Breaths}
When your foe runs the Reforge, do not chase. Weld the penult squares to drown his lanes, or open an SSI−safe Seed of your own while his eyes are on the banner timer.

ℹtem \textbf{Point at the Ford to Close the Sluice}
Verbally threaten a Sanctum seed. Let them waste moves reinforcing it. Then seed from the opposite Sanctum they left under−strengthened.

ℹtem \textbf{Draw the Log from Under the Cauldron}
Do not attack the strong Blue head−on. Attack the weak Red that supports its escape route. Remove the log, and the cauldron tips itself.

ℹtem \textbf{The Cicada Sheds its Shell}
Leave a Red in a seemingly vulnerable position. When they capture it, they reveal the lane your Green uses to sprint for their Home Apex in the endgame.

ℹtem \textbf{Shut the Doors to Catch the Thief}
Do not chase the Reforge runner. Instead, seal every avenue to your own Home Apex. Let them run themselves to exhaustion against walls of your making.

ℹtem \textbf{Feign Madness but Keep Your Balance}
Let your board appear disordered, with gaps and holes. This is not madness. This is a throat, and you are the one who controls when it closes.

ℹtem \textbf{Lure the Tiger Down the Mountain}
Draw their Blue from its fortified Home Apex into the center. The steppe is vaster than their stone keep. On open ground, the hunter has the advantage.

ℹtem \textbf{To Catch Something, First Set it Free}
If you cannot immediately punish a Blue entering your territory, let it pass. Your pieces behind it are a net it does not see. Capture it on its retreat, when its specials are spent and its paths are closed.

ℹtem \textbf{Cast a Stone to Draw the Stallion}
Sacrifice an Orange only when the net is already knotted. A Blue capture is jade if the five breaths of Reforge are already dead on your map.

ℹtem \textbf{Defeat the Enemy by Capturing Their Chief}
The game is not won by capturing pieces. It is won by capturing the Blue and then proving the five breaths of Reforge are a death sentence. All other victories are illusions.

ℹtem \textbf{Let the Rain Fill the Ditch Before You Cross}
Force the enemy to expend effort and resources prematurely; let their strength wane in the waiting.

ℹtem \textbf{The Reed Bends, the Oak Breaks}
Yield to overwhelming force, but do not break; preserve your core to fight another day.

ℹtem \textbf{Weave the Snare with His Own Threads}
Use the enemy's strength, pride, or methods against them; let them build their own trap.

ℹtem \textbf{A Distant Banner Casts No Shadow}
A threat far away is no threat at all; control the center, and distant kings will kneel or starve.

ℹtem \textbf{The Well is Poisoned at its Source}
Strike the enemy's will to fight before striking their body; break their spirit, and their strength follows.

ℹtem \textbf{Do Not Share the Cup of Lies}
Trust, once broken, cannot be mended; if you must deceive, do so cleanly and ensure it is never discovered.

ℹtem \textbf{The Guest Who Outstays His Welcome Steals the Host's Peace}
Allow no force, however small, to remain within your lines uncontrolled; hospitality has its limits.

ℹtem \textbf{A Single Grain Does Not Break the Starving Man's Fast}
Small kindnesses mean nothing to the desperate; offer aid only when it can shift the balance of power.

ℹtem \textbf{Measure the Wind, Not the Banner}
Judge the strength of your enemy not by their display, but by the force that moves them.

ℹtem \textbf{A Silent Horse Leaves No Track}
Move your true strength unseen and unheard; let whispers precede your strike.

ℹtem \textbf{The Forge Glows Hottest Before the Break}
The enemy is most vulnerable not in weakness, but in the moment before their transformation or greatest effort.

ℹtem \textbf{The Vulture Waits Not for the Cry}
True opportunity comes not with fanfare, but in the silence before the fall.

ℹtem \textbf{Speak Softly, and Carry the Unseen Blade}
Let diplomacy smooth the way, but ensure your threat is always present, even if unspoken.

ℹtem \textbf{The Map is Not the Meadow}
Do not become so enamored with your plan that you forget the ground beneath your feet changes with every step.

ℹtem \textbf{A Broken Bridge is Easier to Defend}
Destroy what the enemy relies upon; make their strength a liability.

ℹtem \textbf{Count the Breath Between the Beats}
Victory belongs to he who controls the rhythm; master the pause, and you master the moment.

ℹtem \textbf{The Herd Moves as One, but Feeds as Many}
Unity of purpose does not require equal reward; know who is the shepherd and who is the sheep.

ℹtem \textbf{The Game Ends When the Last Player Forgets the Rules}
True mastery is making the enemy forget they are playing until it is far too late.

\end{enumerate}

\subsection*{0. The Unwritten Strategy: Count Exits, Not Victims}
The greatest strategy is not listed, for it is the breath before all breaths. Before you move, count the number of ways your foe can flee. If the number is greater than zero, you have not yet won. You have only begun to fight.

\section{The Ostrikari (Ikari) −−− Host of the Open Steppe}

\textbf{Confederated riders, oath−law, and the winter mathematics of survival.}\\
The Ostrikari are a people of movement and assembly. In peace, they are wagon−circles, remount strings, and wintering compacts. In war, they become the \emph{Host}: a temporary confederation sworn by oath−rings and overseen by the Thing of Spears. Their strength is not faceless number but the public law that binds them, the herds that feed them, and the memory of winters survived.

\subsection*{Pillars of Identity}
\begin{itemize}
  ℹtem \textbf{Host−Law (Customary & Portable).} Justice travels with the people: wergild (blood−silver), oath−rings, outlawry (``wolf−head''), and the \emph{Thing of Spears} (moot) that can overrule any chief.
  ℹtem \textbf{People of Movement.} Wealth is mouths fed and roads kept open: wagons, remount strings, escort rights, and safe winter pasture.
  ℹtem \textbf{Fostered Bonds.} Children fostered between clans to cool feuds; marriage−compacts braid bloodlines and trade routes.
  ℹtem \textbf{Reputation Economy.} Oathkeepers rise; oathbreakers are unmade. Guest−right kept and promises paid are a leader's true treasury.
\end{itemize}

\subsection*{Why They Did Not Assimilate}
\begin{itemize}
  ℹtem \textbf{Land as Trust, Not Title.} Pasture and wintering grounds held as host commons; fixing them to deeds would shatter reciprocity webs.
  ℹtem \textbf{Law−on−the−Move.} Courts are wagons and circles, not halls of stone; the people would not trade the Thing of Spears for distant writs.
  ℹtem \textbf{Devotion Fit.} Oath, storm, and graves are their holy measures; foreign mandates felt like conquest−of−sky, not covenant.
  ℹtem \textbf{Winter Memory.} A remembered pasture tax that starved a season−−−every clan sings the warning−−−hardens their independence.
\end{itemize}

\subsection*{Law & Custom}
\begin{itemize}
  ℹtem \textbf{Three−Night Guest−Right.} No harm to a guest who keeps house−laws; after three nights, they must pledge or depart. Violation stains the lineage.
  ℹtem \textbf{Blood−Silver (Wergild).} Every wound has a price; pay it, or feud stands. High−born owe more−−−status increases liability.
  ℹtem \textbf{Thing of Spears.} War and law are public; a ring of stood spears may unseat a chief who breaks oath or endangers the winter stores.
  ℹtem \textbf{Outlawry (Wolf−Head).} Outside protection of law; anyone may strike. Declared at dawn by the moot−voice as last resort.
\end{itemize}

\subsection*{Economy Beyond Raiding}
\begin{itemize}
  ℹtem \textbf{Moot−Fairs.} Neutral markets guaranteed by host−law: horses, felt, leather, smoked meats, iron blooms, winter ferry rights.
  ℹtem \textbf{Wintering Compacts.} Fodder and pasture leased from river−lands; paid in escort service and herd−keepers.
  ℹtem \textbf{Craft Lineages.} Pattern−weld smiths, beadmakers, bone−inlay bowyers; artisans travel under \emph{peace−hands} seals.
  ℹtem \textbf{Gift Webs.} Debt and counter−gift braid clans tighter than treaties; ledgers are sung, not inked.
\end{itemize}

\subsection*{War Doctrine of the Host}
\paragraph{Assembly.} Clan warbands + oath−companions + wagon guards. A crisis raises a \emph{war−king} for a season; the Thing must reconfirm after the campaign.

\paragraph{Operational Art.}
\begin{enumerate}
  ℹtem \textbf{Screen & Starve.} Light horse screens, deny forage, scorch the raiding belt the foe needs.
  ℹtem \textbf{Wagon−Lager.} Defensive ring that turns mobility into a fortified camp.
  ℹtem \textbf{Shock Window.} Concentrate for two days: break a gate, force a crossing, or vanish.
\end{enumerate}

\paragraph{Taking & Restraint.} Preference for intimidation, blockade, ransom tables, and bribery; saps and rams appear only with hired or allied engineers. Oathbound limits protect bread−lands and shrines that paid guest−right; violations trigger Thing inquiries.

\subsection*{Faith & Story}
\begin{itemize}
  ℹtem \textbf{Wind & Mound.} Seers keep humming wind−lathes to measure omen−breezes; mound−keepers judge where the host may dwell.
  ℹtem \textbf{Twin Virtues.} \emph{Fury} (storm−oath fulfilled without flinching) and \emph{Measure} (winter mathematics: can we afford this glory?).
\end{itemize}

\subsection*{Factions Within the Confederation}
\begin{itemize}
  ℹtem \textbf{Open−Hand Clans.} Treaty−first, market−forward, eager for federate compacts (escort−for−land).
  ℹtem \textbf{Iron−Hand Clans.} Prestige wars, bride−price politics; insist the host must take before it can trade.
  ℹtem \textbf{Mound−Keepers.} Priestly/judicial lineages; conservative and anti−external−rule, often peacemakers.
  ℹtem \textbf{Young−Riders.} Glory−seeking small bands; can start feuds elders must then cool.
\end{itemize}

\subsection*{Look & Sound}
\begin{itemize}
  ℹtem \textbf{Aesthetic.} Felted cloaks, tablet−woven bands, horsehair and bone fittings, oath−rings on scabbards, shields stitched with storm−signs and ancestor marks, wagons painted with family knots.
  ℹtem \textbf{Music.} Frame drums and throat−hummed hymns; winter feast songs where stanzas are debts−−−who owes, who paid.
\end{itemize}

\subsection*{Mechanical Hooks (Fate's Edge)}
\paragraph{Tags.} \texttt{Mounted}, \texttt{Wagon−Lager}, \texttt{Forage Denial}, \texttt{Reputation Oath}, \texttt{Guest−Right}.

\paragraph{Timers.}
\begin{itemize}
  ℹtem \textbf{Assemble Host [8]} −−− muster complete when filled.
  ℹtem \textbf{Winter Stores [6]} −−− if emptied, Morale drops and Measure overrules Fury.
  ℹtem \textbf{Feud Ignites [4]} −−− missteps push toward blood−debt.
  ℹtem \textbf{External Response [6]} −−− neighbors or distant powers intervene.
\end{itemize}

\paragraph{DV & Position Cues.}
\begin{itemize}
  ℹtem Presenting a foster−token: \textbf{DV −−1} to Parley with elders.
  ℹtem Violating guest−right: immediate \textbf{Position $→$ Desperate} in all social scenes with witnesses; adds +1 tick to \textbf{Feud Ignites}.
  ℹtem Public negotiations at the Thing: truth−telling is \textbf{Dominant}; deceit begins \textbf{Controlled} and flips to \textbf{Desperate} if exposed.
\end{itemize}

\subsection*{Scene Frames}
\begin{itemize}
  ℹtem \textbf{The Wergild Moot.} A death threatens feud; arbitrate price, propose fosterage to end it.
  ℹtem \textbf{Winter Ledger.} The war−king wants a raid; the granary−mother shows the math−−−choose Fury or Measure.
  ℹtem \textbf{Foster Exchange.} Hostages of honor pass between clans; protect the ceremony from spoilers.
  ℹtem \textbf{Oath of Crossing.} Negotiate ferry rights; breaking the oath stains a patron and shuts every gate for a season.
\end{itemize}

\subsection*{Adventure Seeds}
\begin{enumerate}
  ℹtem \textbf{The Wagon−Judge's Bell.} A judge lies slain; find the killer before the Thing outlaws the wrong clan.
  ℹtem \textbf{Host on the Border.} Draft a federate compact; elders demand clauses on guest−right and winter pasture.
  ℹtem \textbf{The Burned Belt.} Bread−lands torched and blame laid falsely; clear the name before the host radicalizes.
  ℹtem \textbf{Storm−King for a Season.} Two chiefs tie the oath−ring to a neutral PC as war−king to stop a feud−−−survive the winter inquiry.
\end{enumerate}

\subsection*{Contrast at the Table}
\begin{itemize}
  ℹtem \textbf{Vilikari:} river−law, road charters, and markets that make bad roads pay by morning.
  ℹtem \textbf{Ostrikari:} host−law, foster bonds, and seasonal assembly into the Host.
  ℹtem \textbf{Ykrul:} route−pricing, breath−count logistics, and proofs over boasts.
\end{itemize}

\begin{fatebox}[Keeper Guidance]
Show mass and momentum through wagons, forage, and remounts−−−not faceless waves. Keep procedural justice visible: \emph{Thing}, wergild, outlawry, fosterage. Let internal arguments breathe: Open−Hand vs Iron−Hand; Fury vs Measure. Make oathkeeping costly and admired−−−villainy is breaking oaths, not being ``barbaric.''
\end{fatebox}

ection{Complications that Feel Like Them}

\begin{itemize}
ℹtem A Gate Cantor switches sides mid−reading−−−because the other side offered repair years for a road you actually use.

ℹtem A Countkeeper discovers the party's favorite clause is forged but correct in spirit. Do you argue fairness or run?

ℹtem The Black Banner arrives to ``help,'' which means your quiet deal now has a drumline.

ℹtem Night Road Heat spikes because a rival used your braid pattern on a crime; clear it by solving their mess in public.
\end{itemize}

\end{document}
