\documentclass[11pt,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ebgaramond}
\usepackage{geometry}
\geometry{a5paper,margin=1.2in,inner=1.5in,outer=0.8in}
\usepackage{graphicx}
\usepackage{epigraph}
\usepackage{lettrine}
\usepackage{xcolor}
\usepackage{verse}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{array}
\usepackage{microtype}
\usepackage{booktabs}

% ----------------------------------------------------------------------
%   Colors
% ----------------------------------------------------------------------
\definecolor{fracture}{RGB}{120, 80, 60}
\definecolor{cracksilver}{RGB}{180, 180, 200}
\definecolor{bloodedge}{RGB}{140, 40, 40}
\definecolor{forestdeep}{RGB}{40, 70, 50}

% ----------------------------------------------------------------------
%   Header / Footer
% ----------------------------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE]{\leftmark}
\fancyhead[RO]{\rightmark}
\fancyfoot[CE,CO]{\thepage}
\fancyfoot[LE,LO]{\textcolor{fracture}{\textit{The Knave's Regret}}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% ----------------------------------------------------------------------
%   Chapter / Section Styles
% ----------------------------------------------------------------------
\titleformat{\chapter}[display]
  {\normalfont\Huge\bfseries\color{fracture}}
  {\chaptername\ \thechapter}
  {20pt}
  {\Huge}
\titlespacing*{\chapter}{0pt}{-50pt}{40pt}

\titleformat{\section}
  {\normalfont\Large\bfseries\color{fracture}}
  {\thesection}{1em}{}

% ----------------------------------------------------------------------
%   Epigraph
% ----------------------------------------------------------------------
\setlength{\epigraphwidth}{0.7\textwidth}
\renewcommand{\epigraphflush}{flushright}
\renewcommand{\sourceflush}{flushright}

% ----------------------------------------------------------------------
%   Custom commands
% ----------------------------------------------------------------------
\newcommand{\scenebreak}{%
  \vspace{1em}%
  \centerline{\textcolor{fracture}{\rule{0.4\textwidth}{0.4pt}}}%
  \vspace{1em}}
\newcommand{\mechanicbox}[1]{%
  \begin{center}\fbox{\parbox{0.9\textwidth}{\textbf{#1}\par\smallskip}}%
  \end{center}}

% ----------------------------------------------------------------------
%   Document start
% ----------------------------------------------------------------------
\begin{document}

% ======================================================================
%   Title page
% ======================================================================
\newgeometry{margin=0.5in}
\begin{titlepage}
  \centering
  \vspace*{2cm}
  {\Huge\bfseries\color{fracture} THE KNAVE'S REGRET\par}
  \vspace{1cm}
  {\Large A Vhasian Expansion for Fate's Edge\par}
  \vspace{1.5cm}
  {\large Set down in the canton of Rouen, by the Silent Stair\par}
  \vspace{2cm}
  {\normalsize \textcolor{fracture}{``When the armor closes, the heart must open---or break.''}}\par
  \vspace{1cm}
  {\small Under the lamp, bread and salt --- and a song for the armor‑bound.\par}
  \vfill
  {\normalsize Year of the Broken Lance, 847 AR\par}
\end{titlepage}
\restoregeometry

\newpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
  \emph{For the knight who unbuckled his gorget to hear a boy's voice.\\
  For the Cantor who carries the silence between notes.\\
  For the chivalry that outlives the peace.}
\end{center}
\vspace*{\fill}
\newpage

% ======================================================================
%   Framing device -- The Cantor's Vigil
% ======================================================================
%--- Prologue -- The Cantor's Vigil -----------------------------------------
\chapter{The Road to Rouen and the Siege}

\section{Scene 1 -- The Orchard Encounter}
The province‑war wind carried ash and the sour scent of burnt orchard trees.  Sir Corin, his mail still slick with the mud of Black Ford, rode through a haze of gray, the weight of twenty years of service pressing against his shoulders like a second skin.  He halted beside a smoldering copse where a boy of twelve knelt over the corpse of a cook, his small fingers working a tarnished ring from the dead man's thumb.

The boy hummed a fragment of a lullaby, the tune uneven as dough under a novice's hands.  He looked up, eyes wide, and the ring slipped free with a soft clink.

``Stand,"" Corin commanded, his voice hoarse from days without water.

The boy obeyed, his shoulders trembling.  The cook's apron, stained with flour, lay abandoned beside the body, a mute testament to the night the cook had fed a starving garrison.

``I should hang you,"" Corin said, meaning it.  He had hanged deserters, looters, men who broke the chivalric code that was the only law left in the fractured duchies.

Instead of begging, the boy began to sing---\emph{the Old Vhasian Ballad} that would later echo in Alain's own voice.

\begin{verse}
\emph{``Forget‑me‑not,''}\\
\emph{The proud knight said,}\\
\emph{Armed and armored for war.}\\
\emph{His lady kissed him through her tears,}\\
\emph{Fearing to see him no more.}\\[4pt]

\emph{Down by the river, where the purple flowers grow,}\\
\emph{He held her close and whispered his love,}\\
\emph{But armed and armored, he felt not her touch---}\\
\emph{The silk of her delicate glove.}\\[4pt]

\emph{She wept. He rose, as if to leave,}\\
\emph{But turned for the flowers instead.}\\
\emph{Overreaching, he tumbled down,}\\
\emph{And as he sank, he said,}\\
\emph{``Forget‑me‑not.''}
\end{verse}

Corin listened, the weight of his armor suddenly apparent---not just steel, but the habit of keeping the world at arm's length.  He realized he had not seen his own face unhelmed for months; he could no longer picture his wife's eyes, only the shape of her letters read by lantern‑light.

``Who taught you that?"" he asked, the edge of the judge softened.

``The cook,"" the boy replied, gesturing to the corpse.  ``He was kind. He taught me to knead, and he taught me the song. Said it was about the weight of things we choose to carry.''

Corin loosened the twine that bound the boy's wrist---just enough to let the blood flow, not enough to free him.

``Walk,"" he said.  ``And sing if you must.  The road to Rouen is long, and the armor grows heavier with every mile.''

The orchard's smoke clung to the low hills like a gray shroud, and the wind carried the distant clatter of siege‑engine wheels as Corin urged Alain onward toward Rouen. Half a league from the smoldering copse, a lone figure emerged from the haze---a knight whose surcoat was faded to the color of old blood, his visor permanently lifted, revealing a face scarred by a dozen small cuts that never seemed to heal. He rode a gaunt courser whose ribs showed through its thin flank, and the beast's eyes were half‑closed, as if the animal, too, had learned to endure without hope.

"Sir Corin of Ethor," the stranger called, his voice flat, devoid of the usual knightly flourish. "I am Sir Garek of the Black Vale. Or of the Western March, depending on which side of last season's quarrel you ask." He gestured at Alain's small frame, then at the boy's clenched fist still clutching the tarnished ring. "You look like you"ve been carrying more than just steel today."

Corin reined in his warhorse, the mail of his hauberk rattling softly. "You know my name. I do not know yours."

Garek shrugged, a motion that seemed to scrape his pauldron against his gorget. "You will, soon enough. Everyone knows the names of men who change sides often enough---they become useful in the telling." He tilted his head toward the smoke rising from Rouen, a thick column now smeared with the orange glow of burning pitch. "The city will fall within the week. Maybe sooner. Lady Isolde of the Western March has already made her bargain with the besiegers: she keeps her lands, they keep the plunder. She calls it pragmatism. I call it the same coin, flipped twice."

He drew a thin, wicked‑looking dagger from his boot and laid it palm‑up on the saddle horn. "Take it. Not for mercy, not for honor---just in case you need to cut something that won"t beg for it. I"ve seen her work before. She doesn"t waste sorrow on the dead. She weighs them like grain."

Corin hesitated, then slipped the dagger into his own belt, feeling its cold weight against his thigh. "You speak as if you pity Rouen."

"Pity?" Garek's laugh was short and dry, like a branch snapping. "I"ve sacked two cities myself---one for the Crown, one against it. Pity is for priests and poets. But I"ll say this: Rouen's people don"t deserve to burn because a baron wanted a higher toll on the bridge. And they don"t deserve to be forgotten after." He glanced at Alain, then back to Corin. "Keep the boy close. The Lady's soldiers won"t ask whose side he's on; they"ll just take his hands for the rings."

He spurred his courser forward, the animal's hooves kicking up ash that settled like forgotten prayers on the road behind him. "Ride on, Sir Corin. May your armor stay tight enough to keep the world out, and loose enough to let you hear the screams you"re trying to silence. I"ll be with the baggage train, counting the silver."

The cynic vanished into the gray, leaving Corin alone with Alain and the weight of the dagger---a reminder that mercy was a luxury, and that some men chose to wear their indifference as openly as their plate.

\section{Scene 2 -- The First Unbuckling (Water for the Wounded)}
The lower town of Rouen was a labyrinth of smoke‑blackened timber and stone, each alley a canyon where the besieging army's trebuchets hurled not only stone but also pots of boiling pitch that hissed as they arced over the walls.  The air tasted of iron and burnt honey; every breath drew in the acrid tang of flaming tar that clung to the cobblestones like a second skin.  Flickering embers drifted down from the inferno above, landing on straw‑filled roofs and igniting them in sudden, orange‑burst blooms that threw grotesque shadows dancing across the faces of the defenders.

Corin moved through this hellscape with Alain tucked close to his side, the boy's small hand clutching the edge of his breastplate for balance.  The citadel's outer bailey was a jagged ruin of shattered stalls---spice sacks torn open, their contents scattering like colored snow over the mud; the scent of cumin and coriander mingled with the ever‑present reek of death.  Militiamen, their faces smeared with soot and blood, manned the breaches, their spears slick with the gore of fallen attackers.

At the courthouse, the magistrate's seals lay broken, the heavy oak doors hanging ajar on corroded hinges.  Inside, the great hall was a cavern of darkness pierced only by the thin shafts of light that slipped through the shattered roof, illuminating dust motes that floated like weary spirits.  Corin's eyes fell upon the writ of absence, the parchment brittle at the edges, the ink faded to a ghostly gray---a stark testament to the abandonment of law in favor of survival.

He read the words aloud, his voice low enough that only Alain could hear: ``By the authority of the absent magistrate, all holdings within the city are forfeit to the defending forces until such time as a legitimate lord may reinstate order."" The boy's shoulders slumped; the realization that even the semblance of justice had dissolved settled over him like a wet cloak.

Corin drew his dagger---not to threaten, but to cut the twine that bound Alain's wrist.  The blade sang a soft, metallic sigh as it severed the fiber, and in the same fluid motion he unfastened the gauntlet on his right hand.  The steel clanged against the stone floor, echoing off the vaulted ceiling like a tolling bell.  The exposed palm was a map of old scars---nicks from practice bouts, a deep gouge from a long‑forgotten duel, and a fresh, thin line where the boy's fingers had once pressed against it.

``Stay,"" Corin murmured, his voice rough from smoke and thirst. ``Help me hold the walls. Help me remember what the armor is for.""

Alain obeyed without hesitation.  He sprinted to the well at the courtyard's edge, the stone lip worn smooth by countless hands.  The water that rose was lukewarm, tinged with the faint rust of the well's iron lining.  He cupped it in his small palms, feeling the liquid slip through his fingers, and hurried to the parapet where a wounded militiaman lay, his thigh pierced by a stray bolt that had punched through the wooden shielding like a vengeful wasp.

As Alain knelt, the water spilled over the man's bandages, turning the linen a dark, mottled red.  The militiaman's eyes fluttered open, pain etched in every line of his face, yet a faint smile touched his lips as the boy's cool touch eased the burning sting.  Corin watched from a few paces away, the weight of his plate pressing against his chest, and felt something inside him shift---a loosening of the rigid code he had clung to like a second skin.  The act of giving water, simple and unarmed, seemed to strip away a sliver of the armor's isolating heft, reminding him that protection could also be an act of openness.

\section{Scene 3 -- The Second Unbuckling (The Burning Building)}
The siege had entered its seventh night, and the sky over Rouen was a roiling cauldron of fire and ash.  Siege engines---massive trebuchets and iron‑clad bombard---threw molten iron balls that screamed as they arced, leaving trails of glowing slag that painted the clouds in sickly orange‑red.  Each impact on the city's walls sent a shudder through the stone, rattling the very foundations and sending showers of sparks raining down like metallic rain.

The lower streets were a furnace.  Timber frames, once sturdy supports for market stalls and homes, now groaned under the heat, their beams twisting and snapping with deafening cracks.  The smell of burning pine mixed with the sharp, metallic tang of overheated iron, creating a perfume that clung to the throat and made each breath feel like inhaling smoke‑filled glass.  Flickering embers floated down like deadly snowflakes, landing on thatched roofs and igniting them in sudden, violent bursts that sent tongues of flame licking upward, devouring everything in their path.

In the midst of this inferno, Alain had slipped away from the clamor of the battlements, his throat raw from singing the old Vhasian ballad that had become his lifeline.  He found a narrow alley where the walls leaned together like exhausted sentinels, their surfaces blackened and slick with soot.  There, a child no older than five lay crammed beneath a heap of soot‑blackened blankets, the tiny chest rising and falling in shallow, frantic bursts.  The boy's face was a mask of ash, his eyes wide and glassy, seeking refuge in any visage that offered even a hint of safety.

Alain scooped the child into his arms, feeling the fragile tremor of the small heart against his forearm.  The infant's eyes fluttered open, locking onto Alain's face with a mixture of terror and a desperate, primal trust.

Just then, a timber beam---once a proud support for a bakery's roof---creaked ominously, its fibers weakened by the relentless heat.  It began to sag, then groaned, and with a sound like a dying beast, it lurched forward toward the alley, its massive end poised to crush everything in its path.

From the smoke emerged Corin, his visor half‑lifted, the darkened metal catching the firelight and throwing it back in jagged shards.  His gauntleted hand shot out, gripping the child's wrist for a heartbeat---a fleeting, almost reverent contact---before he shoved the boy behind a shattered wall and threw his own massive frame across the beam's trajectory.

The impact drove a shock of pain up Corin's spine, his plate absorbing the brunt of the force but transmitting a jarring vibration through every rivet and joint.  The beam struck with a thunderous crash, splintering wood and sending shards flying like deadly shrapnel.  The force jarred Corin's helm loose; the chin strap, strained beyond its limit, snapped with a sharp twang, and the heavy helm clattered to the cobbles, rolling a few feet before coming to rest against a charred wall.

For the first time in years, Corin's face was bare to the night's hellish glow.  Sweat slicked his brow, his eyes---dark as forged iron---scanned the scene, taking in the child's small hand clutching the edge of his breastplate, the flickering flames licking the walls, and the distant cries of the besieged and the besiegers alike.  In that instant, an unspoken apology passed between them: they would not fight side‑by‑side, yet they shared the same sorrow, the same fragile humanity buried beneath steel.

When the beam finally settled, a groan of settling timber rose from the wreckage.  Corin rose, his helm still at his feet, and gathered the child into his arms, cradling the boy against his chestplate.  He carried the wounded infant toward the chapel---the sole sanctuary still standing amidst the rain of arrows and the relentless barrage of fire.  Each step felt heavier, not just from the weight of his plate but from the weight of the choice he had just made: to see, to protect, to be vulnerable in a world that demanded armor as both shield and prison.

\textit{As the last ember‑lit silhouette of the chapel faded behind them, a distant standard snapped against the wind on the ridge beyond the burning town---a sable banner bearing a silver lion, the sigil of the Western March.  The sight was fleeting, swallowed by smoke, but it left a chill in the air that lingered longer than the heat of the flames.}

\section{Scene 4 -- Lady Isolde's Reckoning}
The southern gate, already weakened by days of relentless battering, gave way with a shuddering groan that seemed to echo the very soul of the city.  The invading host poured through the breach like a black tide, their banners snapping in the wind, their faces grim beneath helmets dulled by soot and blood.  The streets became rivers of steel and flame; the clamor of shouted orders, the thud of hooves on broken cobblestones, and the occasional whinny of terrified horses merged into a deafening roar that swallowed the city's usual cadence.

Amidst this chaos, a grim procession emerged from the ruined square: Corin's body, still encased in his battered plate, lay upon a makeshift stretcher hauled by two exhausted soldiers whose own armor was dented and streaked with ash.  The stretcher creaked with each uneven step, the wood protesting under the burden of the fallen knight.

Before the stretcher, Lady Isolde of the Western March halted, her sable cloak heavy with the scent of pine, iron, and the faint, lingering perfume of burnt rosemary that had once garnished the cook's stew---now a cruel reminder of the life that had been snuffed out.  Her gloved hand rested lightly on the breastplate where the insignia of the Order of the Iron Gorget was still faintly visible beneath a film of scorch and soot.  Her eyes, usually sharp as a hawk's talon, were lowered---not in grief but in cold calculation, the kind that measured cost versus gain.

> ``A ransom would have solved this,"" Lady Isolde said, her voice flat as steel, each syllable striking the air like a hammer on an anvil. ``We could have taken a price and avoided the extra loss.""

Around her, the provincial infantry lowered their weapons; the clang of steel fell silent as the reality of the cost settled over the courtyard like a heavy shroud.  Lord Merrick of the Black Vale, who had ordered the sack, watched the scene with a grimace, his jaw tight.  He understood that the price of victory had already been paid in blood, not gold, and that the ledger of war was never balanced by coin alone.

Alain knelt beside the stretcher, his rough‑spun cloak tugging at his shoulders as a gust carried ash and sorrow alike.  He pressed his lips to the cold steel of Corin's breastplate and sang the opening verses of the \emph{Planctus Armatus}, the lament that would become the very soul of the stave.  His voice rose, raw and trembling, over the ruined chapel's cracked arches, weaving a tapestry of sound that seemed to pull the very stones toward a moment of shared humanity.  The soldiers who had come to loot paused, their swords trembling in their hands, as the song wove the image of a knight who chose mercy over steel.

When the final note fell, the chapel's roof shuddered and a fire‑kissed collapse began.  Flames licked the rafters, devouring the ancient timbers with a hungry, crackling fervor.  The wind seemed to echo the boy's forgotten ballad, ``Forget‑me‑not,'' as the stone walls shivered and then gave way, sending a cascade of ember‑laden debris crashing down upon the square.  The air filled with the scent of burning oak and the metallic tang of melting armor, a reminder that even the strongest fortifications could not withstand the relentless hunger of fire.

Thus Sir Corin was carried out, his body a monument to both his courage and his vulnerability, while the conquering lord and lady stood on the threshold of the ruined chapel---recognizing, in their own cold way, that even the most practiced chivalry could not mask the brutal reality of war.


% ======================================================================
%   Table of contents
% ======================================================================
\tableofcontents
\newpage

% ======================================================================
%   Game‑Mechanics Overview (summary of Fate's Edge)
% ======================================================================
\chapter{Game‑Mechanics Overview}
\label{ch:mechanics}

Fate's Edge is a narrative‑first tabletop role‑playing game.  Every action is resolved with the same three‑step loop:

\begin{enumerate}
  \item \textbf{Declare intent \& approach.}  The player says what they are trying to do and which skill they are using.
  \item \textbf{Set Difficulty Value (DV) and Position.}  The GM decides how hard the task is (DV = 2--6) and whether the actor is in a \emph{Dominant}, \emph{Controlled}, or \emph{Desperate} position.  Position modifies the roll and may grant a \emph{Boon}.
  \item \textbf{Roll the dice pool.}  Roll a number of d10 equal to \textbf{Attribute + Skill}.  Each result of 6--10 is a success; each result of 1 generates a \emph{Story Beat (SB)} for the GM.
\end{enumerate}

The \textbf{Outcome Matrix} (see the SRD) determines what happens:

\begin{itemize}
  \item \textbf{Clean Success} -- successes $\geq$ DV and no SBs.  Intent achieved, no complications.
  \item \textbf{Success \& Cost} -- successes $\geq$ DV with one or more SBs.  Intent achieved, but the GM spends the SBs to add a complication, advance a timer, or worsen Position.
  \item \textbf{Partial} -- 0 < successes < DV.  Some progress, a complication, and the player gains 1 Boon.
  \item \textbf{Miss} -- no successes.  No progress, the GM spends SBs for complications, and the player gains 2 Boons.
\end{itemize}

\textbf{Boons} are a player‑controlled resource (max 5, trimmed to 2 at the end of a scene) that can be spent to:

\begin{itemize}
  \item Reroll one die after seeing the pool,
  \item Activate an asset,
  \item Improve Position by one step, or
  \item Pay for a special effect listed in a talent or rite.
\end{itemize}

\textbf{Story Beats (SB)} are the GM's currency for introducing complications, advancing timers, or ticking timers forward.  Every rolled ``1"" creates one SB.

For a detailed description of attributes, skills, talents, timers, and the magic system, see the attached \textit{Fate's Edge -- System Reference Document} (SRD).  The rules presented here are sufficient to play the scenes in this expansion.

\scenebreak

% ======================================================================
%   Introduction -- The Fractured Cantons
% ======================================================================
\chapter{Introduction: The Fractured Cantons}\label{ch:intro}

\epigraph{When the duke rides out, the peasant locks the door.\\
When the duke rides in, the peasant burns the field.\\
This is the chivalry of Vhasia---the armor that protects and isolates in equal measure.}{--- The Gray Wanderer, at the Ethorian marches}

Vhasia is a land of jousts and jurisprudence, of chivalric romance and constant low war.  Since the sundering of the old Crown, the great cantons have fractured into duchies, palatinates, and marcher lordships, each claiming the sunburst seal, each raiding its neighbors for tribute, hostages, and the acknowledgment of lineage.

In the high season the land is peaceful enough for tourneys.  Knights polish their armor until it mirrors the sky.  They ride to the tilt for the favor of ladies whose sleeves they wear, reciting verses of courtly love while the peasantry watches from the lists.  Beneath the romance runs an older current: the Ethorian marches where the forest law still holds, the barrow‑kings who demand tribute in stories, and the sacred groves where armor must be unbuckled or the trees will not witness the oath.

This expansion centers on the tragedy of Sir Corin and the Cantor who witnessed it, but it also expands to encompass the chivalric culture of Vhasia---the disconnect between the armored ideal and the unarmoured heart.

\section{The Nature of Vhasian Chivalry}

The knights of Vhasia follow a code older than the Crown.  It demands:

\begin{itemize}
  \item \textbf{Protection:} To defend the weak, provided they are within the knight's canton.
  \item \textbf{Courtesy:} To honor all ladies, provided their lineage is suitable.
  \item \textbf{Mercy:} To spare the defeated, provided they swear homage.
\end{itemize}

Armor is the physical manifestation of this code---interlocking plates that separate the knight from the world.  To unbuckle the gorget is to risk not only death but intimacy.  Corin's tragedy was that he unbuckled at the wrong moment, for the wrong mercy, and the world was not gentle enough to receive it.

\section{What You Need}
\textbf{Tier:} Seasoned (41--90 XP) \quad \textbf{Players:} 3--6 \quad \textbf{Length:} 3--4 sessions

\subsection{Core Timers}
\begin{itemize}
  \item \textbf{The Armor Closes [6]} --- advances when knights choose protection over mercy, when gorgets are buckled tight, when chivalry becomes siege mentality.  When full, the cantons close their borders and the raids become wars.
  \item \textbf{The Song Escapes [8]} --- advances when Cantors sing the truth, when mercy is shown, when armor is unbuckled.  When full, a moment of connection occurs---though it may be fatal.
  \item \textbf{The Fracture Widens [10]} --- advances with every chivalric gesture that masks a brutal reality.  When full, the Province Wars reignite.
\end{itemize}

% ======================================================================
%   The Cantons of Vhasia
% ======================================================================
\chapter{The Cantons of Vhasia}\label{ch:cantons}

\section{The Ethor Canton}
The loyalist heartland, where the old Crown still holds nominal sway. Its knights are stern, legalistic, devoted to the code. Corin was Ethorian. Their castles are stone, their forests hunted out, their chivalry a matter of record‑keeping.

\section{The Adria Palatinate}
The south western canton, rich in wine and olive groves. Its knights are voluptuaries, lovers of poetry and song. They raid less for land than for hostages to ransom, for the drama of capture and release. Their chivalry is theatrical, performed.

\section{The Ethorian Marches}
The western forests, neighbor to Adria, where the old ways hold. Here, knights must unbuckle their armour to enter the sacred groves. The \emph{Barde de Fer} ballads originate here, warning of the knight who forgets his humanity. The marches are Vhasia's uncivilised heart, dangerous to armoured men.

\section{Istovar}
Across the Belworth from Fenwood, Istovar is held by the Everbloods, cousins to the eponymous royal family of Viterra. Its lords are caught between two loyalties: bound by blood to the Viterran crown, yet entangled by marriage and debt to Vhasian claimants. The castle of Istovar overlooks the river's narrowest crossing, a constant irritant to both realms.

\section{Lence}
The royal seat of Vhasia, sacked by the Fenwoods and their Vhasian Everblood allies under the banner of \textit{royal excess}. They broke the Sun Court, scattered the crown, and then---lacking the strength to hold it---left the throne empty. The crownlands were ravaged, and many barons on both sides of the Belworth have soured on the Fenwoods. Lence remains a half‑ruined city, its courts silent, its bells cracked, its people waiting.

\section{Rouen: The Hub City}
Where the cantons meet and clash. Rouen is a city of inns and tourney fields, of law courts and Ethorian shrines. It is neutral ground by tradition, though neutrality is a luxury of the powerful. The Silent Stair chapel, where Alain sings, is built on the site of Sir Corin's death.
% ======================================================================
%   Factions of the Fracture
% ======================================================================
\chapter{Factions of the Fracture}\label{ch:factions}

\section{The Order of the Iron Gorget}
A knightly order sworn to the ideal of total protection.  They never unbuckle, never eat without helm, never touch skin to skin.  They are fearsome in battle but increasingly isolated from the people they protect.  Corin was once a member; he broke his vow when he spared Alain.

\textbf{Strings:} The Unbroken Seal (a gorget that has never opened), the Weight of Iron (inability to feel light touch), the Ledger of Mercies Refused.

\section{The Cantor's Road}
Not a faction but a calling.  Cantors travel between cantons, singing the news, bearing witness.  They are immune to chivalric protection but subject to its violence.  Alain is the most famous of them, the one who survived Rouen.

\textbf{Strings:} The Cracked Ring, the Second Voice, the Metamorphosis of Witnessing.

\section{The Ethorian Groves}
The old religion of the marches, keeper of the barrow‑kings and the forest law.  They demand that armor be removed on their thresholds.  They remember what the cantons forget.

\textbf{Strings:} The Unbuckled Threshold, the Barrow‑King's Tribute (paid in stories), the Green Seal (which opens only to bare skin).

\section{The Ducal Courts}
Each canton has its court, its heralds, its keepers of chivalric record.  They are the factors of honor, tracking who owes whom the \emph{geste} of protection.

\textbf{Strings:} The Sunburst Seal (fractured), the Tourney List, the Ransom Ledger.

% ======================================================================
%   The Knight's Story (Corin & the Song)
% ======================================================================
\chapter{The Armor and the Song}\label{ch:story}

\section{The Ethorian Ballad of the Iron Bard}
Before Corin there was the legend.  The Ethorians sing of a knight so armored he forgot his own face.  He rode to a grove to meet his lady, but the trees would not part for steel.  He unbuckled, and in that moment of vulnerability he was struck.  He died smiling, having felt the sun.

This is the paradox of Vhasian chivalry: protection requires isolation, but honor requires connection.  The armor is the barrier that makes mercy possible, but it must be lowered to deliver it.

\section{Corin's Choices}
During the siege of Rouen Corin had three opportunities to re‑buck his gorget:

\begin{enumerate}
  \item When he captured Alain (he left it open to hear the song).
  \item When he spared the provincial soldier (he unbuckled to offer water).
  \item When he carried the child (he removed his helm to see the boy's face).
\end{enumerate}

Each unbuckling was a chivalric act.  Each brought him closer to the spear that would end him.

% ======================================================================
%   Patrons
% ======================================================================
\chapter{Patrons of the Armoured}\label{ch:patrons}

\section{Seint Michel, the Judge of the Scar}
Mykkiel is the patron of chivalric law, of the code that binds knights. His Vhasian face is stern: he demands that the code be followed exactly, even when it breaks the heart. Most Vhasians worship the Everflame; he is known as \textbf{Seint Michel}.
Peasants pray to him for fair arbitration; knights swear oaths on his shrines before the joust.

\subsection*{The Oath of the Iron Fastening (Low, 4 XP)}
Seal a promise in armor. It cannot be broken without physical pain. (Mark +1 Obligation.)

\subsection*{The Unbuckled Gorget (Standard, 8 XP)}
Open your armor to show vulnerability. Gain +2 dice to mercy rolls, but --2 to defense until the gorget is closed. (Mark +2 Obligation.)

\section{Seint Isabeau, Who Slumbers Amongst the Lilies}
Ikasha is the secret patron of knights who long to set aside their steel. She offers the dream of connection, the memory of skin. She is dangerous because she makes the armor feel heavy. She is known as \textbf{Seint Isabeau}.

\subsection*{The Forgotten Caress (Low, 4 XP)}
Remember the sensation of rain on skin, of wind in hair. Gain +1 die to resist Corruption, but gain a Corruption tick anyway. (Mark +1 Fatigue.)

\subsection*{The Unhelmed Countenance (Standard, 8 XP)}
Remove your helm in a dangerous place. You are seen truly. The enemy knows your face. (Mark +2 Corruption; gain a String on anyone who witnesses.)

\section{Seint Evroul of the Cairn}
Mor'iraath, the keeper of blood‑debts and ancestor‑witness, walks the Ethorian marches in the guise of a knight in rusted mail, his face hidden by a helm that has never been opened. He is the patron of those who carry the weight of unavenged kin, of oaths sworn on graves, of the cold duty that outlasts love. Most Vhasians fear to speak his name; in the marches, they leave a cup of salt water at the door of the crypt.

\subsection*{The Pledge of the Unquiet Dead (Low, 4 XP)}
Swear before a witness that you will settle a wrong done to your bloodline. Until the wrong is righted, you gain +1 die to all rolls made in pursuit of that vengeance, but you mark 1 Fatigue at dawn each day you do not make progress. (Mark +1 Obligation. This Obligation cannot be cleared except by fulfilling the pledge or by a public act of renunciation that brings shame.)

\subsection*{The Cairn's Toll (Standard, 8 XP)}
Call upon the ancestors to witness a single blow against a foe who has wronged your house. On your next melee attack, you may re‑roll up to two failures. If the blow lands, the foe gains a Condition: \textit{Marked by the Cairn} -- any spirit may locate them for one year. If the blow misses, you suffer 1 Harm (the ancestors' disappointment). (Mark +2 Obligation.)

% ======================================================================
%   New Talents
% ======================================================================
\chapter{The Knight's Legacy: New Talents}\label{ch:talents}

\section{The Unbuckled Gorget (Minor, 4 XP)}
\emph{Requirement: You have shown mercy while vulnerable.}

Once per session you may declare that you remove a piece of armor (helm, gauntlet, or gorget) to show trust.  Gain +1 die to social rolls with honourable enemies, but suffer +1 Harm from the next attack against you.

\section{The Iron Bard's Burden (Major, 8 XP)}
\emph{Requirement: You have forgotten your own face in the mirror of your armor.}

Your Defense is increased by 1, but you cannot regain Stress through social interaction until you unbuckle for a full scene.  The isolation is complete; the protection is total.

\section{The Regret That Remains (Prestige, 12 XP)}
\emph{Requirement: You have survived a moment of connection that destroyed you, as Corin survived until he didn"t.}

You are a living witness to the disconnect.  When you die, you may choose:

\begin{itemize}
  \item Rise as a spirit of the threshold (guardian of a bridge or door, unable to leave but remembering all who pass), or
  \item Pass into the song (become a ballad, sung by Cantors, your name a lesson in mercy).
\end{itemize}

% ======================================================================
%   Campaign Frame
% ======================================================================
\chapter{The Joust and the Raid}\label{ch:campaign}

\section{The Setup}
The party are knights, squires, or witnesses in a Vhasia at peace---but a peace of tightened screws.  The tourney season is beginning.  The raids have not stopped, merely become ritualised.  The Ethorian marches are restive.

\section{The Endings}
\begin{itemize}
  \item \textbf{The Armor Closes:} The party chooses protection over mercy.  The cantons close.  The wars begin.
  \item \textbf{The Song Escapes:} The party unbuckles, shows vulnerability, connects.  They may be destroyed by it, or saved, but the moment is real.
  \item \textbf{The Cantor's Witness:} Alain sings their story.  They become legend---cautionary or celebrated---but they become memory.
\end{itemize}

% ======================================================================
%   Appendix -- The Chivalric Ledger
% ======================================================================
\chapter{Appendix: The Chivalric Ledger}\label{ch:appendix}

\section{The Five Gestes of Vhasian Chivalry}
\begin{enumerate}
  \item The Geste of the Shield (protection offered).
  \item The Geste of the Sword (challenge given).
  \item The Geste of the Helm (courtesy shown).
  \item The Geste of the Unbuckled Gorget (mercy offered).
  \item The Geste of the Song (witness borne).
\end{enumerate}

\section{Ethorian Folklore}
The barrow‑kings demand tribute in stories, not gold.  The sacred groves require bare feet.  The \emph{Barde de Fer} is sung at un‑helming ceremonies, when a knight removes armor for the last time (in death or in retirement).

% ======================================================================
%   Colophon
% ======================================================================
\newpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
\textit{This expansion was written in the Chapel of the Silent Stair, on paper made from tourney programmes.  The ink is lamp‑black and the oil of unbuckled armor.  The binding is felt, to muffle the bells.}\\[1em]

\textit{The armor is polished.  The song is sung.  The gorget is open.}\\[2em]

--- The Clerk of Rouen, in the year of the joust
\end{center}
\vspace*{\fill}

\chapter*{Appendix B: The Fractured Crown -- A Political Primer for the Cantons of Vhasia}
\addcontentsline{toc}{chapter}{Appendix B: The Fractured Crown}

\section*{A Note from the Raven Road}
\addcontentsline{toc}{section}{A Note from the Raven Road}

\textit{Set down by Scribe Valerius of the Thepyrgos Synod, at the request of the Raven Road. An addendum to "The Knave's Regret," with apologies for the lack of poetry.}

A decade ago, Duke Braedon Fenwood of Viterra crossed the Belworth with fire and iron. His target was not the Vhasian people, but the Sun Court itself. He sacked Lence, scattered the royal family, and---most curiously---left the throne empty. "A crown that must be seized is not a crown," he is said to have told his knights. "Let them fight over the pieces. It costs me less than a garrison."

The result is a power vacuum that has turned Vhasia into a chessboard of claimants, condottieri, and cousins. Every duke, baron, and bastard with a drop of Sun blood now eyes the throne. Meanwhile, the common folk pay the tithes of three masters: the local lord, the claimant they hope will win, and the Vritann tax collector who never left.

This appendix provides tools for Game Masters to run the political chaos of Vhasia: faction timers, claim mechanics, cross‑border entanglements, and a parlement mini‑game for those who prefer legal dagger‑work to steel. It is designed to be used alongside \textit{Blood on the Belworth} and the core \textit{Political Intrigue} expansion.

\section*{The Vacant Throne Timer [10]}
\addcontentsline{toc}{section}{The Vacant Throne Timer }

This campaign timer represents the race to fill the power vacuum in Lence. It ticks forward whenever:

\begin{itemize}
    \item A claimant gains a significant advantage (a major battle, a noble defection, a foreign alliance).
    \item The party brokers a truce or exposes a claimant's weakness (ticks backward).
    \item The GM spends 2+ Story Beats on off‑screen political manoeuvring.
\end{itemize}

When the timer fills, a claimant (or a coalition) takes Lence and declares themselves king. The party's subsequent adventures will be shaped by whether they backed the winner, opposed them, or tried to stay neutral.

\section*{Claims as Strings and Assets}
\addcontentsline{toc}{section}{Claims as Strings and Assets}

A \textbf{Claim} is a legal or blood right to a title, land, or influence. In Vhasia, a claim is also a weapon.

\paragraph{Claim Strings}
A minor claim (e.g., "third cousin of the last queen," "a debt of honor from a Fenwood raid") functions as a String. Once per session, the holder may invoke the claim to gain \(+1\) die on a social roll involving a noble, a court, or a border dispute. The String is consumed after use.

\paragraph{Claim Assets}
A major claim (e.g., "the Sunburst Seal in my possession," "the loyalty of the Ethorian Marches") is a Minor or Standard Asset (4 XP or 8 XP). Its free off‑screen effect might be: "Once per session, declare that a minor noble refuses a demand because your claim is superior." Its Boon activation might allow you to force a duel of honor or call a moot.

\paragraph{The Sunburst Seal (Major Asset, 12 XP)}
The physical seal of the old crown, lost in the sack of Lence. Whoever possesses it is considered the legitimate king by those who respect such symbols. Effects: free off‑screen -- "Your herald is granted audience at any court in Vhasia"; Boon activation -- "You declare a law that stands until challenged by a higher authority (i.e., a claimant with more military force)." The Seal also draws envy; the GM may start a \textit{Claimant's Hunt} [6] timer when it is used.

\section*{Factions of the Fracture}
\addcontentsline{toc}{section}{Factions of the Fracture}

\begin{longtable}{|p{0.22\textwidth}|p{0.22\textwidth}|p{0.22\textwidth}|p{0.22\textwidth}|}
\hline
\textbf{Faction} & \textbf{Goal} & \textbf{Clock} & \textbf{Strings / Leverage} \\
\hline
\endfirsthead
\hline
\textbf{Faction} & \textbf{Goal} & \textbf{Clock} & \textbf{Strings / Leverage} \\
\hline
\textit{The Sun‑in‑Exile} & Restore the rightful & Restoration [8] -- advances & "The Queen's Ring," "A wet‑nurse's \\
(surviving royal family) & bloodline & when they gain a military & testimony," "The lost register of births" \\
& & backer or a border lord's oath & \\
\hline
\textit{The Fenwood Claim} & Install a Viterran puppet & Puppet's Progress [6] -- advances & "Letters of safe‑conduct from Viterra," \\
(Duke Braedon's nephew, Touénot l'Argent) & (or himself) & when Touénot l'Argent wins a tourney, & "The Belworth Bridge tolls," \\
& & a Parlement vote, or a skirmish & "A hostage oath from a border baron" \\
\hline
\textit{The Ducal Compact} & Maintain the vacuum & Compact's Cohesion [6] -- advances & "A sealed apology," "A marriage contract," \\
(three powerful dukes) & while building their & when they agree on a common & "A map of shared salt‑pans" \\
& own power & enemy; retreats when they squabble & \\
\hline
\textit{The Condotta Companies} & Get paid, or carve out & Payday [8] -- advances each & "A captain's gambling debt," \\
(mercenary captains) & their own fief & season they are not paid; when & "A banner stolen from a rival," \\
& & full, a company deserts or flips & "A letter of marque from a phantom king" \\
\hline
\end{longtable}

\section*{The Armor Closes -- A Chivalric Timer }
\addcontentsline{toc}{section}{The Armor Closes -- A Chivalric Timer }

This timer (size 6) tracks the emotional and psychological isolation of Vhasian knighthood. It advances when knights choose protection over mercy, when gorgets are buckled tight, when chivalry becomes siege mentality. When full, the cantons close their borders and the raids become wars. The GM may also spend 2 SB to advance this timer when a character refuses to show vulnerability.

The companion timer, \textbf{The Song Escapes} (size 8), advances when Cantors sing the truth, when mercy is shown, when armor is unbuckled. When full, a moment of genuine connection occurs -- though it may be fatal.

\section*{The Belworth Border -- Dual Loyalties and Claims}
\addcontentsline{toc}{section}{The Belworth Border -- Dual Loyalties and Claims}

Many Vhasian lords hold lands on both sides of the Belworth, a legacy of the old united kingdom. This creates constant friction: \textbf{dual loyalties} and \textbf{overlapping claims}.

\begin{itemize}
    \item \textbf{Dual Loyalty Timer } -- A border baron who is also a Viterran count might have a timer [6]. When it advances, he leans toward his Viterran liege; when it retreats, toward his Vhasian claimant. Players can influence this timer by gifts, threats, or exposing secrets.
    \item \textbf{Fenwood's Shadow} -- Duke Braedon Fenwood does not openly press a claim, but his agents are everywhere. A \textit{Fenwood Intrigue} [8] timer ticks when the party crosses him. When full, a Fenwood assassin or a legal writ from Viterra arrives.
    \item \textbf{The Belworth Bridge} -- This key crossing is a source of constant toll disputes. A \textit{Bridge Toll} [4] timer ticks each time the party uses it without a proper pass. When full, the bridge is closed, forcing a dangerous ford or a long detour.
\end{itemize}

\paragraph{Touénot l'Argent (NPC)}
\textit{"They call me Touched Not the Silver. The silver I refused at birth -- the silver I will not take as a bribe -- the silver that will buy my crown."}

The nephew of a distant Fenwood cousin, Touénot was raised in a Vhasian hedge‑court, educated by a renegade Cantor, and knighted on a battlefield where he refused the customary silver price for his first prisoner. Now he styles himself the \textit{Heir by Abstention} -- the one man who never owed a Fenwood debt.
x
\textbf{Stats (Tier III):} Presence 4, Wits 3, Body 3, Spirit 3. Skills: Command 3, Sway 4, Melee 2, Lore 2 (Vhasian precedents).

\textbf{Strings:} "A silver thorn (his token)," "The testimony of a Lantern‑Rite witness," "A promise from the River Syndicate."

\textbf{Clock:} \textit{Touénot's Coup} [6] -- advances when he wins a Parlement vote, a joust, or a skirmish. When full, he marches on Lence with a banner of silver and black.

\textbf{Adventure Hook:} Touénot offers the party a charter to collect the "Silver Tithe" -- a tax on every crossing of the Belworth Bridge. The charter is legal, but the previous collector was found floating face‑down, a silver coin in his mouth. The ghost now haunts the bridge, whispering: \textit{"Touénot... l"Argent... the price..."}

\section*{The Parlement Mini‑Game (Legal Duel)}
\addcontentsline{toc}{section}{The Parlement Mini‑Game (Legal Duel)}

When claimants cannot settle a dispute by steel, they turn to the Parlement of Lence -- a rump assembly of clerics, lawyers, and surviving magistrates who still meet in a half‑ruined hall. The party may be called as witnesses, advocates, or judges.

\textbf{Procedure (3 beats):}

\begin{enumerate}
    \item \textbf{Precedent} -- Each side presents a written citation (roll Wits + Lore, DV 4). Winner gains the "Precedent Edge" (+1 die in the final ruling).
    \item \textbf{Witness} -- Each side calls a witness (real or bribed). The GM sets a social DV (3‑5) based on the witness's loyalty. Winner gains "Witness Favor" (can force a re‑roll on one die in the final ruling).
    \item \textbf{Judgment} -- The presiding magistrate (or the party, if they have the authority) declares a verdict. The player designated as judge rolls Spirit + Command (DV 5). Success means their preferred outcome holds; failure means the magistrate rules against them, or the losing side rejects the verdict and the dispute escalates to combat.
\end{enumerate}

\noindent
\textit{Outcomes:} A legal victory grants a Claim String on the defeated party, and may advance or retreat a faction timer.

\section*{New Talents for Vhasian Knights and Courtiers}
\addcontentsline{toc}{section}{New Talents for Vhasian Knights and Courtiers}

\paragraph{Claimant's Voice (Minor, 4 XP)}
You have trained in the art of the Parlement. Once per session, when you invoke a Claim String, you may also force a single NPC to make a Resolve test (DV 4) or reveal their true allegiance in that matter.

\paragraph{Banner of the Empty Throne (Major, 8 XP)}
You carry a standard that claims to represent the true king (even if you do not know who that is). Once per session, you may declare a safe‑conduct for yourself and your allies in any Vhasian canton. The safe‑conduct is honoured until you are exposed as a fraud (GM discretion) or a rival claimant denounces you.

\paragraph{The Unbuckled Gorget (Minor, 4 XP)}
\textit{Requirement: You have shown mercy while vulnerable.} Once per session you may declare that you remove a piece of armor (helm, gauntlet, or gorget) to show trust. Gain \(+1\) die to social rolls with honourable enemies, but suffer \(+1\) Harm from the next attack against you.

\paragraph{The Iron Bard's Burden (Major, 8 XP)}
\textit{Requirement: You have forgotten your own face in the mirror of your armor.} Your Defense is increased by 1, but you cannot regain Stress through social interaction until you unbuckle for a full scene. The isolation is complete; the protection is total.

\paragraph{The Unbuckled Gorge (Prestige, 10 XP, requires \textit{Unbuckled Gorget})}
When you remove a piece of armour to show vulnerability, you also gain the ability to speak one true sentence that cannot be interrupted. The sentence must be a single clause, and it is heard by all present as if you had whispered it in their ear. Afterward, you mark 1 Corruption (the weight of unarmoured truth). \textit{"I was afraid."} --- Ser Roran.

\section*{Sample Claimant: Touénot l'Argent du Istovar (NPC)}
\addcontentsline{toc}{section}{Sample Claimant: Touénot l'Argent Fenwood (NPC)}

The nephew of Duke Braedon, Touénot l'Argent was raised in Vhasia as a hostage and now styles himself the "Heir by Deliverance." He is handsome, charismatic, and utterly ruthless -- a mirror of Richard III before the fall.

\textbf{Stats (Tier III):} Presence 4, Wits 3, Body 3, Spirit 2. Skills: Command 3, Sway 3, Melee 2, Deception 3.

\textbf{Strings:} "A promise from the Ethor Duke," "A letter of safe‑conduct from Viterra," "The captain of the Lence garrison (bribed)."

\textbf{Clock:} \textit{Touénot l'Argent's Coup} [6] -- advances when he wins a Parlement session, a joust, or a skirmish. When full, he marches on Lence. The party may accelerate or decelerate this timer.

\textbf{Adventure Hook:} Touénot l'Argent offers the party a charter to collect tolls on the Belworth Bridge in exchange for their support. The charter is genuine -- but the previous toll‑collector was murdered, and his ghost now haunts the bridge, demanding to know who betrayed him.

\section*{Blood on the Belworth -- A Campaign Frame}
\addcontentsline{toc}{section}{Blood on the Belworth -- A Campaign Frame}

The adventure \textit{Blood on the Belworth} (from the \textit{Political Intrigue} expansion) is designed to run in this exact period. Its core mechanics -- Situation Timers, Counter‑Timers, Faction Sheets, and the Hedge‑Law Moot -- fit seamlessly with the Vhasian setting.

GMs are encouraged to:

\begin{itemize}
    \item Use the \textit{Armor Closes} timer as a backdrop to the knighthood scenes.
    \item Allow the \textit{Unbuckled Gorget} talent to grant a bonus in the hedge‑law scenes (showing vulnerability before the Lantern‑Rites).
    \item Introduce the \textit{Sunburst Seal} as a potential prize for resolving the ferry murders.
    \item Let players who invoke a \textit{Claim String} automatically know one of the clue threads in Act II.
\end{itemize}

 % ----------------------------------------------------------------------
%   BRETONIAN TOURNEY RULES -- THE ROUND OF FIVE VIRTUES
%   For Fate's Edge: a chivalric judging system inspired by Arthurian
%   romances, the Matter of France, and the code of courtly love.
% ----------------------------------------------------------------------

\section*{The Round of Five Virtues}
\addcontentsline{toc}{section}{The Round of Five Virtues}

In Bretonian tourneys, victory is not decided by blood alone. A panel of nobles, heralds, and a chosen \textit{Lady of Honor} judges each passage of arms by five knightly virtues. A combatant may be unhorsed yet praised; another may strike true yet be disgraced.

\subsection*{The Five Virtues}
Each exchange is scored from 0 to 2 in five categories:

\begin{description}
    \item[Courage] -- Facing danger without recklessness, standing firm when pressed.
    \item[Courtesy] -- Respect toward opponent, the crowd, and the Lady of Honor.
    \item[Mercy] -- Sparing the fallen, offering aid, refusing to strike a helpless foe.
    \item[Prowess] -- Skill with lance, sword, or axe; horsemanship; tactical acumen.
    \item[Truth] -- Fighting under one's true banner, keeping sworn oaths, honesty in the lists.
\end{description}

\subsection*{The Judging Panel}
\begin{itemize}
    \item \textbf{Lady of Honor} -- presides, offers a favor (scarf, sleeve, or ring) to a chosen competitor. Her word breaks ties.
    \item \textbf{Herald of the Lists} -- records blows, enforces rules, announces scores.
    \item \textbf{Two Knight- Adjudicators} -- one chosen by each competing house or party.
    \item \textbf{(Optional) Cleric or Druid} -- for tourneys blessed by the Old Faith or the Church.
\end{itemize}

\subsection*{Scoring Rounds}
A formal tourney is fought in three passes (joust) or three bouts (melee). Between rounds the judges confer openly. Use the scorecard below.

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Virtue} & \textbf{0 Points} & \textbf{1 Point} & \textbf{2 Points} \\
\hline
Courage & Fleeing, cowardice, refusing combat & Holding ground, steady defense & Charging a superior foe, protecting an ally \\
\hline
Courtesy & Insults, ignoring the Lady, foul language & Proper salute, respectful address & Public praise of opponent, kneeling to the Lady \\
\hline
Mercy & Striking a fallen foe, refusing quarter & Accepting surrender, allowing recovery & Offering aid to wounded enemy, freeing a captive \\
\hline
Prowess & Missed strikes, broken lance, unhorsed early & Clean hit, solid defense, staying mounted & Three lances splintered, disarming the foe, unhorsing \\
\hline
Truth & False colors, broken oath, concealed weapon & Fighting under own banner, truthful declaration & Refusing an unchivalrous advantage, confessing an error \\
\hline
\end{tabular}
\caption{Bretonian Tourney Scorecard -- The Five Virtues}
\end{table}

\subsection*{Special Rules}

\paragraph{The Lady's Favor}
Before the tourney, each competitor may request a favor from a noble lady (not necessarily the Lady of Honor). A favor grants +1 die to one Prowess roll but imposes -1 Position on any Courtesy roll (the crowd watches closely). Favors are physical tokens (sleeve, ribbon, ring) that can be displayed or returned.

\paragraph{The Pledge of Truth}
All combatants must swear the \textit{Pledge of Truth} before the assembled court. Breaking the pledge (e.g., fighting under false colors) immediately loses 2 Virtue points in Truth and may disqualify the offender. The oath is:
\begin{quote}
\emph{``By my spurs, by my liege, and by the light of this Lady's eyes, I swear I bear no false heraldry, no poison, and no enchantment save that which a knight may honourably carry. May my horse falter if I lie; may my sword turn in my hand if I deceive.''}
\end{quote}

\paragraph{The Judgment of the Twelve}
For disputes that cannot be settled by the panel, any knight may demand the \textit{Judgment of the Twelve} -- six champions a side, fighting a mounted melee until one party yields or the Lady of Honor declares the truth revealed. The Judgment ignores normal scoring; the outcome is accepted as divine witness.
\paragraph{Honor & Shame Tokens}
Competitors earn \textbf{Honor Tokens} for acts of mercy, courtesy, or truth (spend to improve Position once). They earn \textbf{Shame Tokens} for cowardice, cruelty, or lies (GM may spend to worsen Position). Three Shame Tokens in a single tourney lead to public disgrace -- the knight is stripped of spurs and cannot compete again without a pilgrimage.

\subsection*{Sample Scorecard (Joust -- Three Rounds)}
Use this table at the table.

\noindent
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Round} & \textbf{Sir Alaric} & \textbf{Sir Renaud} & \textbf{Notes} \\
\hline
Courage (1st pass) & 1 & 2 & Renaud charges into broken lance \\
Courtesy (1st pass) & 2 & 1 & Alaric salutes the Lady, Renaud mutters \\
Mercy (1st pass) & 2 & 2 & Both pull up to avoid trampling a page \\
Prowess (1st pass) & 1 & 2 & Renaud unhorses Alaric \\
Truth (1st pass) & 2 & 2 & Both banners true \\
\hline
Courage (2nd pass) & 2 & 1 & Alaric remounts under fire \\
... & ... & ... & ... \\
\hline
\textbf{TOTAL} & \textbf{14} & \textbf{12} & \textbf{Alaric wins by Virtue} \\
\hline
\end{tabular}

\subsection*{Integration with Fate's Edge}
\begin{itemize}
    \item Use the \textbf{Seven Bell Court} mechanics for any generic formal duel (Form, Spirit, Intent). Use \textbf{Round of Five Virtues} for Arthurian / Bretonian / crusader settings.
    \item Timers such as \textit{Crowd Mood}, \textit{Tournament Integrity}, and \textit{Lady's Patience} work unchanged.
    \item The Lady of Honor may be a PC with appropriate social talents (e.g., \textit{Courtly Graces}, \textit{Silver Tongue}).
\end{itemize}

\subsection*{GM Guidance}
Bretonian tourneys are as much about reputation as about victory. Encourage players to narrate their shields, their favors, and their vows. Use the Virtues to reward roleplay -- a knight who dismounts to help a fallen foe should gain Mercy points even if he loses the match.

A simple mantra: \emph{``The lance is judged, but the heart is weighed.''}

\section*{Closing Notes from the Raven Road}
\addcontentsline{toc}{section}{Closing Notes from the Raven Road}

\textit{The crown is a story we tell ourselves to make the violence bearable. The claimants tell competing versions. The people listen and pay their tithes. And the Fenwood duke sits across the river, sharpening his quill.}

\textit{Do not mistake the vacuum for peace. It is only the pause between breaths.}

\vspace{1em}
\noindent
--- Dusana of the Raven Road \\
\textit{On the Belworth ferry, waiting for the fog to lift}

\input{copyright}
\end{document}