```latex
\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{lmodern}
\usepackage{multicol}

\geometry{margin=1in}
\setstretch{1.1}
\allsectionsfont{\normalfont\sffamily\bfseries}
\pagestyle{fancy}
\fancyhf{}
\rhead{\thepage}
\lhead{Lich's Dungeon of Penance}
\renewcommand{\headrulewidth}{0.4pt}

\titleformat{\section}{\Large\bfseries\sffamily}{\thesection}{1em}{}
\titleformat{\subsection}{\large\bfseries\sffamily}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\bfseries\sffamily}{\thesubsubsection}{1em}{}

\newcommand{\sectionbreak}{\clearpage}

\begin{document}

\begin{center}
{\Huge \textbf{Lich's Dungeon of Penance}} \\
\vspace{0.5cm}
{\Large A tragic, choice-driven dungeon crawl for Fate's Edge}
\end{center}

\section*{How to Run This Adventure}

\subsection*{Floor-as-Leg Procedure}
\begin{enumerate}[label=\arabic*.]
    \item \textbf{Seed the Floor:} Treat each floor as a Travel "leg." Draw your seed (Spade = place, Heart = actor, Club = pressure, Diamond = leverage) from a reskinned "Dungeon Deck," or choose elements that match the listed floor. Highest rank sets the Floor Clock.
    \item \textbf{Set Position:} Start at Controlled. Adjust during play for bold risks, insight, or rising backlash.
    \item \textbf{Track the Floor Clock [4–6]:} Use [4] for compact scenes, [5] for tense puzzles/social play, [6] for complex combats or multifaceted dilemmas.
    \item \textbf{SB Economy = Revelation Engine:} Prefer spending SB to reveal memory, raise stakes, or tilt loyalties over raw damage. Draw from the Deck of Consequences when the scene earns it (multiple 1s, spikes in tension) to color omens, debts, or emotional fallout.
    \item \textbf{End with a Choice:} Each floor ends on a fork that reframes the lich. Choices tune Position for the next floor and unlock boons or burdens.
\end{enumerate}

\subsection*{Deck Usage in the Dungeon}
\begin{itemize}
    \item \textbf{Hearts:} Shame, love, vows, memory-ghosts.
    \item \textbf{Clubs:} Breakage, traps, relentless pursuers.
    \item \textbf{Diamonds:} Wards, keys, rights of passage, bargains.
    \item \textbf{Spades:} Necromancy, omens, soul-anchors, whispers.
\end{itemize}

\subsection*{XP \& Boons}
\begin{itemize}
    \item Award XP for discoveries, hard choices, embracing costs, and major objectives.
    \item Grant Boons for partials/misses and compassionate play; allow Boons to steady Position or reroll at pivotal moments.
\end{itemize}

\subsection*{Final Persuasion Clock}
\begin{itemize}
    \item The last floor is resolved via a Persuasion Clock [6–8] with three destinations: Continue Penance, Find Peace, Join the Living. Table and player choices across prior floors shift starting Position and DV. See the Final Floor at the end of the document.
\end{itemize}

\subsection*{Tone \& Safety}
\begin{itemize}
    \item This adventure centralizes grief, broken vows, and contrition. Check lines/veils, use safety tools, and invite players to define what reconciliation looks like in this world.
\end{itemize}

\section*{Tiering \& Difficulty Dials}
\begin{itemize}
    \item \textbf{Party Tier I–II:} Default clocks [4–5]; sparse reinforcements; SB favors revelations and single-issue conditions.
    \item \textbf{Tier III+:} More [6] clocks, layered wards, and factional crossfire; SB can call reinforcements, gear failure, or scene shifts in addition to revelations.
\end{itemize}

\section*{Entry: The Forbidden Library (Hook)}
A silent archive under municipal seal; brittle chains anchor a violet door. Research exposes this dungeon's purpose: a penitential labyrinth. Breaking the seal opens to the Landing Stair and a breeze like the last breath of a vow.

\textbf{Skill Hooks:} Lore (necromantic jurisprudence), Insight (contrition vs. cruelty), Tinker (seal-hardware), Arcana (sympathy lines).

\textbf{Stakes:} You descend to hold the lich accountable—or to free him.

\section*{The 30 Floors of Penance}
Each floor below lists: Memory, Environment, Clock, Handles, SB Spends, and Choices. Start Position is Controlled unless stated.

\subsection*{1) The Abandoned Apprentice}
\begin{itemize}
    \item \textbf{Memory:} He left his first student locked out of the rite to claim credit.
    \item \textbf{Environment:} Workshop split by a glowing chalk-circle; tools hum with stolen charge.
    \item \textbf{Clock:} [5]. Handles: Arcana (counter-glyph), Sway (validate the apprentice-echo), Tinker (ground the circle), Skirmish (shatter constructs).
    \item \textbf{SB Spends:} (1) The apprentice's notes accuse a small theft; (3) Ward backlash binds a PC's shadow.
    \item \textbf{Choices:} Restore Credit (gain Boons +1; Position +1 next floor), Break the Circle (fast, risks backlash), Erase the Notes (Diamond omen: debt marked).
\end{itemize}

\subsection*{2) The Broken Vow}
\begin{itemize}
    \item \textbf{Memory:} He abandoned his wedding mid-ritual for discovery.
    \item \textbf{Environment:} An aisle of frost, a bride of falling petals, guests frozen mid-gasp.
    \item \textbf{Clock:} [6] Wedding Disintegration (space fractures into cold shards).
    \item \textbf{Handles:} Presence/Sway (honor the vow), Faith/Rites (bind a surrogate rite), Wits (spot the escape sigil), Duel (hold off jealous guardians).
    \item \textbf{SB Spends:} (2) Choir turns accusatory; (4) The ring binds a PC to a test of fidelity.
    \item \textbf{Choices:} Honor (blessing vs undead), Reject (curse of chill—Desperate start next floor), Third Path (promise a different covenant; gain Omen).
\end{itemize}

\subsection*{3) The Fallen Comrade}
\begin{itemize}
    \item \textbf{Memory:} He fled, leaving a shield-mate to die.
    \item \textbf{Environment:} Sand-choked battlefield; banners as nooses.
    \item \textbf{Clock:} [5]. Handles: Command (rally ghost cohort), Medicine (salvage the dying echo), Skirmish (cut through memory-bandits), Move (dodge sand-wights).
    \item \textbf{SB Spends:} (1) Confusion fog; (3) Old campaign clock ticks—reinforcements arrive.
    \item \textbf{Choices:} Stay \& Hold (Position +1), Rescue \& Retreat (Boons +1, but a ghost follows), Leave (Diamond: debt to the dead).
\end{itemize}

\subsection*{4) The Rejected Love}
\begin{itemize}
    \item \textbf{Memory:} He spurned mortal affection to pursue immortality.
    \item \textbf{Environment:} Night garden of silver lilies; a heart in a glass cloche beats slowly.
    \item \textbf{Clock:} [4]. Handles: Sway (sincere apology), Theft (return a token), Arcana (unbind cloche), Performance (dance the pledge back).
    \item \textbf{SB Spends:} (2) Flowers whisper your worst doubt; (3) The cloche draws breath from the room.
    \item \textbf{Choices:} Return the Token (gain Favor; +1d once later), Break \& Bury (end the beat; add Gloom tag next floor), Promise to Carry (take a vow-Condition to clear Gloom).
\end{itemize}

\subsection*{5) The Betrayed Mentor}
\begin{itemize}
    \item \textbf{Memory:} He stole forbidden pages.
    \item \textbf{Environment:} Spiral study; every step recites footnotes; the mentor's cane taps from above.
    \item \textbf{Clock:} [5]. Handles: Deception (stage a confession), Tinker (swap facsimile pages), Arcana (trace original ink), Skirmish (duel the cane).
    \item \textbf{SB Spends:} (1) A marginalia-imp bites at evidence; (3) A secret page flares, blinding.
    \item \textbf{Choices:} Return Pages (Position +1), Keep Knowledge (gain Technique, but GM banks a Debt), Expose Hypocrisy (mentor-echo breaks; Consequences Deck draw now).
\end{itemize}

\subsection*{6) The Children of the Plague-House}
\begin{itemize}
    \item \textbf{Memory:} He quarantined too harshly, dooming survivors.
    \item \textbf{Environment:} Boarded infirmary, lullaby through vents.
    \item \textbf{Clock:} [6]. Handles: Medicine (triage echoes), Insight (weigh utilitarian cost), Rites (quiet passing), Move (thread contagion wards).
    \item \textbf{SB Spends:} (2) Cough spreads to a PC (Fatigue), (4) A screaming vent opens—time loss.
    \item \textbf{Choices:} Mercy (gain Blessing of Rest), Iron Law (Desperate next floor; but gain Efficiency boon later), Risk It All (save two echoes; take Debt).
\end{itemize}

\subsection*{7) The Crown Informant}
\begin{itemize}
    \item \textbf{Memory:} He traded a friend to the authorities.
    \item \textbf{Environment:} Tea-house murmurs; lacquered masks tilt toward you.
    \item \textbf{Clock:} [4]. Handles: Deception (cover story), Sway (reframe as lesser betrayal), Streetwise (flip the informant chain), Skirmish (mask-enforcers).
    \item \textbf{SB Spends:} (1) Friend-echo recognizes you; (3) A sealed writ lands—law binds choices.
    \item \textbf{Choices:} Confess (gain Forgiveness token; spend later to ease a social DV), Double Down (gain leverage; group marks Stain), Cut the Chain (burn the writ; Diamond omen tracks you).
\end{itemize}

\subsection*{8) The Oath of the Doorwarden}
\begin{itemize}
    \item \textbf{Memory:} He failed a sacred watch.
    \item \textbf{Environment:} Temple gate, seven locks, one missing tooth.
    \item \textbf{Clock:} [5]. Handles: Tinker (forge tooth), Rites (reconsecrate), Endure (hold the door), Shoot (pin the breach-creature).
    \item \textbf{SB Spends:} (2) The hinge screams; (3) A pilgrim-echo tries to push through.
    \item \textbf{Choices:} Stand the Watch (Position +1 next two floors), Open to All (gain Popular Favor; Club draw now), Bar Forever (close one shortcut; reduce later DV by 1 once).
\end{itemize}

\subsection*{9) The Scholar's Duel}
\begin{itemize}
    \item \textbf{Memory:} He discredited a rival through lies.
    \item \textbf{Environment:} Lecture-hall, chalk specters, a bell that tolls on falsehoods.
    \item \textbf{Clock:} [4]. Handles: Evidence (reconstruct data), Sway (track to apology), Deception (plant a correction), Skirmish (ink-spirits).
    \item \textbf{SB Spends:} (1) Note-mites chew proofs; (3) The bell tolls wrong—everyone loses a turn.
    \item \textbf{Choices:} Publish Retraction (Boons +1), Win Again (gain Repute, but a critic haunt follows), Bind the Bell (stabilize truth; Diamond licenses your re-entry).
\end{itemize}

\subsection*{10) The Siege of Hearth-Row}
\begin{itemize}
    \item \textbf{Memory:} He sold a neighborhood to fund research.
    \item \textbf{Environment:} Narrow lanes; house-gods glare from soot.
    \item \textbf{Clock:} [6]. Handles: Command (evacuate echoes), Trade (find funds), Skirmish (block enforcers), Rites (invoke hearth-right).
    \item \textbf{SB Spends:} (2) Roof collapse; (4) A hostage ledger appears.
    \item \textbf{Choices:} Buy Back Time (lose coin/asset; +2 segments cleared), Name Names (Position -1; gain Coin), Break the Ledger (boon: Hearth's Favor; draw Spade omen).
\end{itemize}

\subsection*{11) The Pact of the Dry Well}
\begin{itemize}
    \item \textbf{Memory:} He broke a village's water-pact.
    \item \textbf{Environment:} Salt rings, hung jars, something sighs below.
    \item \textbf{Clock:} [5]. Handles: Rites (appease spirit), Survival (re-route flow), Tinker (raise bucket engine), Sway (mediate terms).
    \item \textbf{SB Spends:} (1) Cracks spread underfoot; (3) A thirst-curse licks your lips.
    \item \textbf{Choices:} Patch the Pact (gain Token: Water-right), Take the Depths (immediate aid; mark Obligation-style Condition), Walk Away (Desperate start next floor).
\end{itemize}

\subsection*{12) The Unburied}
\begin{itemize}
    \item \textbf{Memory:} He let enemies lie unburied to study them.
    \item \textbf{Environment:} Windmill of bones; chorused flies say your name.
    \item \textbf{Clock:} [4]. Handles: Rites (burial), Arcana (sever tethers), Skirmish (put down bonewights), Endure (carry the weight).
    \item \textbf{SB Spends:} (2) Bone-sail cleaves cover; (3) A skull remembers your childhood.
    \item \textbf{Choices:} Bury (Blessing vs Death magic), Catalogue (+1 effect vs undead later; gain Stain), Confession (reduce next DV by 1 once).
\end{itemize}

\subsection*{13) The Rival's Heart}
\begin{itemize}
    \item \textbf{Memory:} He stole a rival's lover through glamour.
    \item \textbf{Environment:} Ballroom of mirrors; your reflections accuse.
    \item \textbf{Clock:} [5]. Handles: Performance (dance the truth), Sway (public apology), Arcana (break glamour), Skirmish (mirror-doppel fights).
    \item \textbf{SB Spends:} (2) Floor becomes slick with spilled secrets; (4) Mirror eats a die from a pool.
    \item \textbf{Choices:} Undo Glamour (Position +1), Keep the Lie (gain Favor now; lose a friend-later), Break the Mirrors (Spade omen: shard sticks in soul).
\end{itemize}

\subsection*{14) The Broken Oath to the Dead}
\begin{itemize}
    \item \textbf{Memory:} He promised to guard a grave.
    \item \textbf{Environment:} Hill with standing stones; an honored name is scratched out.
    \item \textbf{Clock:} [4]. Handles: Survival (restore cairn), Rites (oath-binding), Skirmish (grave-robbers), Insight (why he failed).
    \item \textbf{SB Spends:} (1) Cold wind steals breath; (3) A name tries to climb into a PC's mouth.
    \item \textbf{Choices:} Restore Name (gain Ancestor's Watch boon), Seal It Away (safer path opens; lose that boon), Let the Dead Speak (learn truth; draw Deck now).
\end{itemize}

\subsection*{15) The Research Fire}
\begin{itemize}
    \item \textbf{Memory:} He set a fire to hide evidence.
    \item \textbf{Environment:} Stacks blaze; soot-spirits chant.
    \item \textbf{Clock:} [6]. Handles: Move (create firebreak), Tinker (sprinkler glyphs), Skirmish (ash wraiths), Insight (find the real truth he buried).
    \item \textbf{SB Spends:} (2) Roof beam falls; (4) A "saved" book is false.
    \item \textbf{Choices:} Save Lives (Blessing), Save Proof (unlock clue elsewhere; mark Guilt), Let it Burn (Position -1; but gain Power cipher later).
\end{itemize}

\subsection*{16) The Orphan's Blade}
\begin{itemize}
    \item \textbf{Memory:} He trained a child assassin.
    \item \textbf{Environment:} Rooftop training ground; bell-lanterns swing.
    \item \textbf{Clock:} [5]. Handles: Sway (deprogram), Skirmish (disarm gently), Acrobatics (catch the fall), Arcana (unbind kill-phrase).
    \item \textbf{SB Spends:} (2) Knife flashes at a throat; (3) Kill-phrase echoes in the bells.
    \item \textbf{Choices:} Break the Phrase (gain Guarded Word talisman), Finish the Lesson (gain deadly ally; mark Stain), Adopt Responsibility (downtime obligation reduces future DV once).
\end{itemize}

\subsection*{17) The Siege Engine Test}
\begin{itemize}
    \item \textbf{Memory:} He field-tested a weapon on a village.
    \item \textbf{Environment:} Abandoned ballista field; stones weep.
    \item \textbf{Clock:} [6]. Handles: Tinker (sabotage), Command (evacuate), Skirmish (fight engineers), Evidence (expose sponsor).
    \item \textbf{SB Spends:} (2) Misfire blasts the party; (4) Sponsor-echo offers coin to walk away.
    \item \textbf{Choices:} Sabotage (lower later harm), Expose (gain Leverage; attract enemies), Take the Coin (resources now; Position -1 next two floors).
\end{itemize}

\subsection*{18) The University Vote}
\begin{itemize}
    \item \textbf{Memory:} He blocked reparations to a harmed cohort.
    \item \textbf{Environment:} Senate in candle-gloom; ballots bleed wax.
    \item \textbf{Clock:} [4]. Handles: Sway (coalition-build), Deception (parliamentary trick), Law (craft binding motion), Skirmish (duel a bully-doctor).
    \item \textbf{SB Spends:} (1) Whip flips votes; (3) Clock gains a segment as rules shift.
    \item \textbf{Choices:} Pass Restitution (gain Public Grace), Table It (reduce heat; lose Grace), Burn the Charter (scene shift now; future boon vs bureaucracy).
\end{itemize}

\subsection*{19) The Expedition}
\begin{itemize}
    \item \textbf{Memory:} He misled peers into a lethal dig.
    \item \textbf{Environment:} Wind-scoured necropolis digsite; tent pegs creak.
    \item \textbf{Clock:} [6]. Handles: Survival (safe path), Arcana (read tomb script), Skirmish (sand-ghuls), Tinker (shore trenches).
    \item \textbf{SB Spends:} (2) A trench collapses; (4) Ancient plague-sigil coughs.
    \item \textbf{Choices:} Seal the Tomb (Blessing vs miasma), Plunder (gain Relic; mark Curse), Name the Dead (lower Persuasion DV later).
\end{itemize}

\subsection*{20) The Confessor's Lament}
\begin{itemize}
    \item \textbf{Memory:} He used a lover's confession in a paper.
    \item \textbf{Environment:} Confessional of glass; ink drips inside.
    \item \textbf{Clock:} [4]. Handles: Presence (apology), Evidence (destroy the paper), Deception (convince the echo it was mutual), Arcana (blank the confession).
    \item \textbf{SB Spends:} (2) Ink stains a PC's hands (Stain), (3) The echo asks a PC to confess something true.
    \item \textbf{Choices:} Public Apology (gain Truth's Edge die once), Quiet Erasure (safer now; debt later), Trade Truths (everyone tells one; Boons +1).
\end{itemize}

\subsection*{21) The Coven Gate}
\begin{itemize}
    \item \textbf{Memory:} He stole a ritual and sold it.
    \item \textbf{Environment:} Standing stones with borrowed sigils, a buyer waiting in fog.
    \item \textbf{Clock:} [5]. Handles: Arcana (restore rightful words), Sway (renegotiate), Skirmish (stop the exchange), Theft (swap the grimoire).
    \item \textbf{SB Spends:} (2) Buyer names an ally; (4) Stone groans; threshold tears.
    \item \textbf{Choices:} Return It (blessing vs wards), Sell Again (gain Coin; draw Spade omen), Bind the Gate (close a later hazard).
\end{itemize}

\subsection*{22) The City That Starved}
\begin{itemize}
    \item \textbf{Memory:} He speculated on grain during famine.
    \item \textbf{Environment:} Market of empty baskets; scales hang lopsided.
    \item \textbf{Clock:} [6]. Handles: Trade (unlock grain), Command (food line order), Skirmish (stop mob), Rites (invoke charity rite).
    \item \textbf{SB Spends:} (2) Spoilage spreads; (4) A grain-storm blinds.
    \item \textbf{Choices:} Release Stores (Public Grace), Protect the Warehouse (resources; mark Stain), Name the Brokers (reduce faction heat later; draw Diamond omen now).
\end{itemize}

\subsection*{23) The High Stair}
\begin{itemize}
    \item \textbf{Memory:} He let a rival fall to claim a seat.
    \item \textbf{Environment:} Endless stair with two rails; one is greased.
    \item \textbf{Clock:} [4]. Handles: Move (save-the-rival), Sway (admit the truth), Skirmish (cut the saboteur cord), Tinker (fix the rail).
    \item \textbf{SB Spends:} (1) Your boot slips; (3) A third climber grabs your ankle.
    \item \textbf{Choices:} Lift Them (Position +1), Admit It (DV -1 later), Step Over (gain Seat token; Desperate start next floor).
\end{itemize}

\subsection*{24) The Dragon's Loan}
\begin{itemize}
    \item \textbf{Memory:} He took a predatory loan that doomed a village.
    \item \textbf{Environment:} Coin-hall with a sleeping drake bound in ink.
    \item \textbf{Clock:} [5]. Handles: Trade (renegotiate), Arcana (smudge sigil), Skirmish (fight collectors), Evidence (prove coercion).
    \item \textbf{SB Spends:} (2) Interest compounds: add a segment; (4) Draconic omen opens its eye.
    \item \textbf{Choices:} Refinance to Mercy (lower future Diamond DVs), Pay in Blood (harm now; later grace), Burn the Ledger (scene shift; draw Deck now).
\end{itemize}

\subsection*{25) The Faith He Broke}
\begin{itemize}
    \item \textbf{Memory:} He abandoned a patron's calling mid-rite.
    \item \textbf{Environment:} Chapel half-lit; icons inverted.
    \item \textbf{Clock:} [4]. Handles: Rites (renew vow), Sway (argue theology), Endure (fast), Skirmish (exorcise dissenters).
    \item \textbf{SB Spends:} (1) Omen turns its face; (3) The icon drinks a PC's breath.
    \item \textbf{Choices:} Renew (gain Omen: Steady), Apostasy (cut ties; Spade omen follows), Name a New Path (unique boon keyed to table's Patron themes).
\end{itemize}

\subsection*{26) The Ward Unkept}
\begin{itemize}
    \item \textbf{Memory:} He ignored a binding to chase a rumor.
    \item \textbf{Environment:} House with mirrored halls; a captive shade counting.
    \item \textbf{Clock:} [5]. Handles: Arcana (rebinding), Insight (why he left), Tinker (replace anchor), Skirmish (contain spillover).
    \item \textbf{SB Spends:} (2) Counting accelerates; (4) Two mirrors disagree (Position -1).
    \item \textbf{Choices:} Finish the Binding (boon vs ghosts), Free the Shade (ally later; draw Deck), Break All Mirrors (scene safer; lose ally).
\end{itemize}

\subsection*{27) The Pilgrim's Road}
\begin{itemize}
    \item \textbf{Memory:} He extorted pilgrims for passage.
    \item \textbf{Environment:} Gatehouse with tally-rope and oath-bucket.
    \item \textbf{Clock:} [4]. Handles: Law (revoke toll), Sway (refund ritual), Trade (pay now, leverage later), Skirmish (stand-down guards).
    \item \textbf{SB Spends:} (1) A saint's bell rings; (3) A riot starts.
    \item \textbf{Choices:} Refund (Public Grace), Keep Tolls (resources; Stain), Turn the Gate (open safe shortcut later).
\end{itemize}

\subsection*{28) The Crown of Ashes}
\begin{itemize}
    \item \textbf{Memory:} He accepted a crown promised to another.
    \item \textbf{Environment:} Throne room under drifting ash, two crowns on one chair.
    \item \textbf{Clock:} [6]. Handles: Sway (abdicate), Duel (trial-by-steel), Evidence (prove the promise), Arcana (split the crown).
    \item \textbf{SB Spends:} (2) Ash-snow blinds; (4) Spectral court demands new oaths.
    \item \textbf{Choices:} Abdicate (DV -1 later vs authority), Rule by Fear (Desperate tag sticks 1 floor), Share Crown (unlock ally later).
\end{itemize}

\subsection*{29) The Night of Names}
\begin{itemize}
    \item \textbf{Memory:} He erased a persecuted lineage to hide himself.
    \item \textbf{Environment:} Scriptorium where names fade as you read.
    \item \textbf{Clock:} [5]. Handles: Evidence (restore records), Sway (admit lineage theft), Skirmish (stop inquisitors), Arcana (fix name-magic).
    \item \textbf{SB Spends:} (2) Your name blurs; (4) A PC risks losing their past.
    \item \textbf{Choices:} Restore Line, Burn It All (safety now, haunting later), Sign Your Own True Name (gain Anchor boon).
\end{itemize}

\subsection*{30) The Last Light}
\begin{itemize}
    \item \textbf{Memory:} He chose lichdom to pay for everything.
    \item \textbf{Environment:} A quiet study of candles. The lich sits, unarmed.
    \item \textbf{Clock:} Persuasion [6–8]. Start Position depends on your prior mercy and confession floors.
    \item \textbf{Handles:} Sway/Presence (argue), Insight (understand what he pays), Evidence (lay out the ledger of harms and reparations), Rites (open a door beyond).
    \item \textbf{SB Spends:} (2) Candles gutter images of those you saved; (4) A beloved echo asks you to forgive him—or forbid forgiveness.
    \item \textbf{Endings (Pick by Persuasion Result):}
    \begin{itemize}
        \item \textbf{Continue Penance:} The lich seals himself deeper. Grant Fetter's Boon (1/session: erase 1 SB from a roll; GM banks +1).
        \item \textbf{Find Peace:} He passes on; a Soul-Light remains (1/arc: remove a Desperate → Controlled shift for one scene).
        \item \textbf{Join You:} He returns in frail flesh; a Mortal Oath binds him to aid you against a defined threat. Mark a Debt clock that ticks on misuse.
    \end{itemize}
\end{itemize}

\section*{Running Notes \& Conversions}
\begin{itemize}
    \item \textbf{Scaling Clocks:} If the group is breezing through, upgrade [4] → [5] or add a Ward/Alert sub-clock keyed to a floor's trap.
    \item \textbf{Position Drift:} Mercy, confession, and restitution raise Position next floor; cruelty or evasion drop it. Keep a simple toggle per PC.
    \item \textbf{SB Spend Menu (Dungeon Flavor)}
    \begin{itemize}
        \item 1 SB: Memory whispers; +1 Supply segment; evidence goes missing.
        \item 2 SB: Alarm raised; lose cover; a lesser guardian arrives; vow-echo asks a costly question.
        \item 3 SB: Reinforcements; key gear breaks; rail tick to a larger tragedy.
        \item 4+ SB: Trap springs; authority/omen arrives; scene shifts to harsher truth.
    \end{itemize}
    \item \textbf{Boons as Compassion:} When players console, bury, confess, or repair—hand out Boons. Let them convert Boons to stabilize Position at moral cruxes.
    \item \textbf{Deck Cadence:} Aim for 1–2 Deck draws per session unless the table wants chaos. Translate results into fiction: omens, debts, vows—not just penalties.
\end{itemize}

\section*{Prep Shortcuts}
\begin{itemize}
    \item \textbf{Make a Name Ledger:} three columns—Harmed, Helped, Unresolved. Move names between columns as choices land.
    \item \textbf{Keep a Shard Kit:} mirror shard, ring, ledger scrap, bone-petal—physical tokens that move between floors to track themes.
    \item \textbf{Sketch the lich in three moods:} Proud, Weary, Resolved. Check one at the end of each session; bring that mask to Floor 30.
\end{itemize}

\section*{Rewards \& Advancement}
\begin{itemize}
    \item This adventure is a major arc: award arc XP at resolution. Each floor can also grant small, flavorful boons (temporary tags, one-use rites, favors). Compassionate resolutions often trade immediate safety for long-term strength—lean into it.
\end{itemize}

\section*{Appendix: Quick Floor Hooks by Pillar}
\begin{itemize}
    \item \textbf{Combat-forward:} 3, 10, 12, 15, 17, 28
    \item \textbf{Social-forward:} 2, 4, 7, 13, 18, 30
    \item \textbf{Puzzle/Investigation:} 1, 5, 8, 11, 19, 29
\end{itemize}

Use as a one-shot (pick 6–8 floors), a short arc (12–16 floors), or the full penitence (all 30). The lich isn't "defeated"—he's convinced.

\section*{Final Floor — The Last Light (revised)}
\begin{itemize}
    \item \textbf{Memory:} He chose lichdom to pay for everything—and to be unable to stop paying.
    \item \textbf{Environment:} A quiet study of candles. Thirty flames, each the height of a floor you climbed. The lich sits unarmed; the air is bright with names unsaid.
    \item \textbf{Start Position:} Controlled if you showed mercy/restitution on ≥10 floors; Risky otherwise. Any violent approach starts Desperate.
\end{itemize}

\textbf{Optimal Resolution:} Nonviolent
\begin{itemize}
    \item \textbf{Primary Clock:} Persuasion [6–8] (6 for merciful runs, 8 for mixed).
    \item \textbf{Handles:} Sway/Presence (argue the end of penance), Insight (name what he fears to lose), Evidence (lay the ledger), Rites (open a door beyond).
    \item \textbf{Name-Litany Bonus:} For each unique Key Name recovered this run, mark +1 progress on the Persuasion Clock (max +6 from names).
    \item \textbf{Auto-Succeed — Recite the Thirty:} If the party recites all 30 true names, in order or as a coherent litany, the Persuasion Clock immediately completes.
\end{itemize}

\textbf{SB Spends (nonviolent):}
\begin{itemize}
    \item 2—Candles gutter images of those you saved.
    \item 3—A beloved echo begs you to forbid forgiveness.
    \item 4—Scene shift: you argue inside one of the memories; stakes double.
\end{itemize}

\textbf{The Thirty-Spoken (Name-Litany Rite)}
\begin{itemize}
    \item \textbf{Requirements:} All thirty Key Names, plus at least one Shard Kit token from three different floors.
    \item \textbf{Test:} One PC leads the litany (Rites or Presence, DV 2–3). Others Aid by offering tokens or short remembrances.
    \item \textbf{Risk:} Mispronunciations/omissions tick Memory Collapse [4] (fractured scene; +1 DV until mended).
    \item \textbf{Effect:} On success, the lich yields and accepts your judgment (choose an ending). On failure, you still fill segments equal to correctly spoken names before the error.
\end{itemize}

\textbf{Violent Resolution (Daring, Suboptimal)}
\begin{itemize}
    \item \textbf{Unfettered Lich (Phase Boss)}
    \item \textbf{Wards:} Immune to mundane harm; harm only lands in an exchange where a PC invokes a true Key Name tied to a past wrong.
    \item \textbf{Phylactery:} The unlit thirty-first candle. Find with Insight (what isn't grieving) or Arcana (trace sympathy), DV 3.
    \item \textbf{SB Spends (violent):} 1—Soul-fire lashes (scorch harm). 2—Gear failure. 3—Memory collapse (split-arena). 4—Cataclysmic ward (choose: harm 2, or fall/lose footing and -1 Position next scene).
    \item \textbf{Kill Condition:} Reduce him to cinders and snuff the thirty-first candle while speaking any true Key Name. Otherwise he reforms at dusk.
    \item \textbf{Cost:} Violent victory denies the full blessing of the relic and marks Stain on the Name Ledger; merciful echoes turn against you.
\end{itemize}

\textbf{Endings}
\begin{itemize}
    \item \textbf{Continue Penance:} He seals himself deeper. Gain Fetter's Boon (1/session erase 1 SB from a roll; GM banks +1 SB). The stair closes for a year and a day.
    \item \textbf{Find Peace:} He passes onward; a Soul-Light remains (1/arc: shift Desperate → Controlled for one scene).
    \item \textbf{Join You:} He returns in frail flesh; a Mortal Oath binds him against a named greater threat. Track Debt [4] that ticks on misuse.
\end{itemize}

\textbf{Treasure—Only if The Thirty-Spoken:} When the party recites all 30 names and he yields, he bequeaths the Glaive of the Thirty (below). If he's slain, only a fine (non-enchanted) glaive lies here unless later awakened by returning at least ten names to the living record.

\textbf{Relic: Glaive of the Thirty (Enchanted)}

A two-handed polearm of pale iron, its edge etched with faint, incomplete names that brighten when spoken with care.
\begin{itemize}
    \item \textbf{Tags:} reach • two-handed • solemn • name-bound
    \item \textbf{Attunement:} Fully awakens only if the party recited all 30 names at the final floor. Otherwise it functions as masterwork steel.
\end{itemize}

\textbf{Boons}
\begin{itemize}
    \item \textbf{Litany Edge (1/session):} Speak a true name from your Name Ledger while striking or standing guard → steady Position one step or +1 effect vs undead, oath-spirits, bound shades.
    \item \textbf{Mercy Carried Forward (1/arc):} When you choose a nonviolent exit in a scene of grievous harm, clear 1 segment from an active Doom/Alert/Ward sub-clock.
    \item \textbf{Witness's Reach (passive):} When defending the powerless or honoring the harmed, gain +1d on the first relevant action of the scene.
\end{itemize}

\textbf{Costs \& Limits}
\begin{itemize}
    \item \textbf{Name Hunger:} Each use of Litany Edge consumes one distinct name for this arc (no repeats until next arc).
    \item \textbf{Contrition Bound:} If you cut down a pleading foe without offering terms, mark Stain; the glaive goes dormant until you restore a name to the world (public restitution, grave restored, ledger repaired).
\end{itemize}

\textbf{Awakening After Violence}
\begin{itemize}
    \item If claimed through combat, the relic may awaken after the party returns 10+ names to the living record. When it stirs, each bearer feels a brief weight on the shoulder—like a hand asking for gentleness.
\end{itemize}

\section*{The Penitent Lich — Runekeeper Patron}

A Patron of names kept, vows witnessed, and debts made visible. Power flows through confession, restitution, and the binding force of true names.

\textbf{What the Penitent Lich Is}

Once mortal, now an anchor of memory. He refuses oblivion until every name he harmed is spoken correctly and set in its rightful place. He grants power to those who lift names, keep watches, and choose mercy first.

\textbf{Portfolio:} true names • oaths • remembrance • wards • restitution \\
\textbf{Symbols:} a guttering candle; a ledger ribbon; a ring that won't warm; a mirror shard. \\
\textbf{Favored Deeds:} return a name to the world; bury the unburied; witness a vow without centering yourself; refuse an easy cruelty.

\subsection*{Tenets (Keep These)}
\begin{enumerate}
    \item Name the Harmed: When you act, speak who is owed—not what you want.
    \item Mercy Before Might: Try a nonviolent avenue before you draw steel or spell.
    \item Watches Are Kept, Not Owned: If you take a ward, keep it—or find a worthy keeper.
    \item Debt Is Not Silence: Confess, repair, and let the record show it.
    \item No Erasures: Do not profit by deleting names, histories, or lineages.
\end{enumerate}

\subsection*{Runekeeper: What You Get}

When you swear to the Penitent Lich as Runekeeper, you learn to inscribe small, portable bindings called runes. Each rune is a single-use inscription prepared during downtime or a quiet scene.
\begin{itemize}
    \item \textbf{Preparing Runes:} In downtime, prepare 2 runes from your list; carry at most 3 at a time. You may overwrite an unspent rune by marking Obligation to the Patron.
    \item \textbf{Invoking Runes:} Invoking a rune is usually Controlled and takes a quick action unless the scene is in chaos (then Risky).
    \item \textbf{Interacting with Position/DV:} Runes usually raise Position one step or give +1 effect toward their stated sphere.
\end{itemize}

\subsection*{Devotions \& Boons}

Track your standing with the Penitent Lich as Favor [0–5]. Increase Favor by naming the harmed, choosing restitution, keeping hard watches, or completing a rite of remembrance. Decrease Favor by erasing names, refusing confession, or striking down a pleading foe without terms.

\textbf{Favor 0 — Initiate of the Ledger}
\begin{itemize}
    \item \textbf{Patron Gift — Rune:} Mark of Witness: Chalk or ash mark that steadies Position once per scene when you defend the powerless or truth.
    \item \textbf{Prayer—Soft Candle:} Once per session, dim a scene's hostility: -1 segment from an active Doom/Alert/Ward sub-clock after you speak who is being protected.
\end{itemize}

\textbf{Favor 1 — Keeper of the Door}
\begin{itemize}
    \item \textbf{Rune:} Threshold: Place on a doorway/line. For one scene, allies crossing it get +1d to resist separation or ambush.
    \item \textbf{Boon—Confessor's Ears:} In a social exchange where you admit a fault, gain +1d to read motive (Insight-type handles) this scene.
\end{itemize}

\textbf{Favor 2 — Binder of Names}
\begin{itemize}
    \item \textbf{Rune:} True Name: Inscribe a true name on cloth/metal. For one action that targets the named (aid or oppose), choose: +1 effect or steady Position. If used to harm, mark Stain unless you offered terms first.
    \item \textbf{Boon—Ledger Glimpse:} Once per session, ask the GM: "What would restitution look like here?" Gain a concrete, actionable step.
\end{itemize}

\textbf{Favor 3 — Warden of Restitution}
\begin{itemize}
    \item \textbf{Rune:} Redress: Break this seal to clear 2 segments from a Debt/Obligation clock you owe—but you must immediately perform a small, public act of repair.
    \item \textbf{Boon—Vigil:} When you keep watch over a warded thing for a full scene, everyone on your side begins the next scene one Position higher.
\end{itemize}

\textbf{Favor 4 — Litany-Bearer}
\begin{itemize}
    \item \textbf{Rune:} Litany: Speak a short list of rightful names. For the next minute, your side treats the scene's first Desperate roll as Controlled. Consumes one distinct name from your Name Ledger.
    \item \textbf{Boon—Mercy Carried Forward:} When you choose a nonviolent exit in a grievous scene, remove 1 segment from a Memory/Alert clock.
\end{itemize}

\textbf{Favor 5 — Keeper of the Last Light}
\begin{itemize}
    \item \textbf{Rite:} The Thirty-Spoken: If your Ledger holds 30 true names from a single penitential place, you may resolve a confrontation with persuasion instead of violence; on success, also refresh one spent rune.
    \item \textbf{Boon—Absolution's Edge:} Once per arc, declare immunity to fear/charm for your party in a scene where you speak for the harmed.
\end{itemize}

\subsection*{Runes (Runekeeper List)}

Pick from these when you prepare:
\begin{itemize}
    \item Witness: Mark a surface; your side gets +1 effect to protect witnesses or evidence this scene.
    \item Threshold: See Favor 1.
    \item True Name: See Favor 2.
    \item Redress: See Favor 3.
    \item Peace: Anyone who voluntarily sheaths a weapon within arm's reach of this rune gains +1d to their next social action in this scene.
    \item Rest: Set by a body or grave; undead entering must test (DV 2). On fail, they hesitate and reveal who they were.
    \item Atonement: Place on an object tied to harm; returning it to its rightful keeper steadies Position for your group and creates a minor Boons token.
\end{itemize}

\subsection*{Rites \& Invocations}
\begin{itemize}
    \item \textbf{Candle of Remembrance (scene):} Light a small candle and speak a harmed name. For this scene, you may re-roll 1s once on a social or investigation action that centers that person.
    \item \textbf{Seal of Safe Passage (scene):} Trace a circle and recite three names you have helped. For the next crossing (gate, bridge, alley), reduce the first Ward/Alert sub-clock you trigger by 1.
    \item \textbf{Confessor's Vigil (downtime):} Keep watch at a threshold named in your Ledger. Clear 1 stress/strain and prepare 1 extra rune next downtime.
\end{itemize}

\subsection*{Compels \& Taboos}

\textbf{Compel (GM Tool):} Offer a boon or clear 1 segment of an impending trouble if the Runekeeper attempts a nonviolent option first. Refusing marks Stain or loses 1 Favor (player's choice). \\
\textbf{Taboos (Break = lose 1–2 Favor):} Erase a name for gain; profit from falsified records; abandon a sworn watch without successor; kill a pleading foe without terms.

\subsection*{Offerings \& Obligations}
\begin{itemize}
    \item Restore a headstone or ledger page.
    \item Return a stolen token (ring, book, oath-stone).
    \item Keep a night's watch for someone who cannot.
    \item Publicly confess a relevant fault and make a concrete repair.
\end{itemize}

Mark Obligation [4] when you take aid without an offering. Fulfilling a listed offering clears 2 segments.

\subsection*{Synergy: Glaive of the Thirty}

If your party earned the Glaive of the Thirty, you may speak a Ledger name while guarding or striking to gain +1 effect or steady Position (once/scene). Re-using the same name before the next arc marks Stain.

\subsection*{Using This Patron in the Lich's Dungeon}
\begin{itemize}
    \item \textbf{Name Ledger:} Keep three columns: Harmed / Helped / Unresolved. Moving names from Unresolved → Helped is how you organically earn Favor.
    \item \textbf{SB Spending Guidance:} Spend SB to reveal truths, tighten oaths, or summon witnesses, not just to spike harm.
    \item \textbf{Nonviolent First:} With this Patron, players should feel tangible benefits for choosing mercy—Position bump, DV hints, and clock relief.
\end{itemize}

\subsection*{Example Invocations}
\begin{itemize}
    \item "By the breath that warmed ALYSSA WHITE-LILY's ring, let no vow be mocked while this mark holds."
    \item "Threshold, keep—the harmed may pass before the hungry."
    \item "Ledger, open—show me what restitution looks like, not what is easy."
\end{itemize}

\subsection*{GM Notes (Quick)}
\begin{itemize}
    \item Favor should move. Reward visible repair. Make taboos costly but recoverable through clean acts.
    \item When in doubt, let a rune raise Position or grant +1 effect toward mercy, truth, or watchkeeping rather than raw damage.
\end{itemize}

\section*{Penitential Corruption — Tables \& Rules (Fate's Edge)}

Use this when the dungeon's necromancy stains the soul, when names are erased, or when the party spurns mercy. It integrates with Position, DV, SB, the Deck of Consequences, and the Name Ledger.

\subsection*{How Corruption Works}
\begin{itemize}
    \item \textbf{When to Roll:} Gain Minor Corruption on small cruelties (breaking etiquette of remembrance, exploiting grief, petty desecrations). Gain Major Corruption for taboos (erasing a name, killing a pleading foe without terms, shattering a ward, stealing a ritual for profit, phylactery desecration without rites).
    \item \textbf{Track:} Each PC keeps a Corruption [0–6] track. Each roll adds +1 Corruption first, then you roll on the relevant table. Some results add more.
    \item \textbf{Stain vs. Corruption:} Stain is a narrative tag that can clear through play; Corruption is a mechanical track with thresholds below.
    \item \textbf{GM SB Hook:} In scenes of cruelty or erasure, the GM may spend 2 SB to force a Minor Corruption roll (once/scene).
    \item \textbf{Name Ledger Hook:} At Corruption ≥3, you cannot move any name from Unresolved → Helped without public restitution.
\end{itemize}

\subsection*{Thresholds}
\begin{itemize}
    \item 3+ Corruption: Choose a Mark (visible stigma). Deception about your past suffers DV +1. You cannot exceed Favor 3 with the Penitent Lich until you reduce to ≤2.
    \item 6 Corruption: You are Oath-Bound Revenant until cleansing (see below): all violent actions start Desperate; Glaive of the Thirty (if held) goes dormant; a personal Haunt follows you as a GM tool.
\end{itemize}

\subsection*{Cleansing Corruption (pick/stack as fiction allows)}
\begin{itemize}
    \item Public Confession + Concrete Restitution: -1
    \item Bury the Unburied / Restore a Headstone or Ledger Page: -1
    \item Keep a Named Vigil (one full scene) at a relevant threshold: -1
    \item Return a Stolen Token to its rightful keeper: -1
    \item Restore a Lineage / Undo an Erasure: -2
    \item The Thirty-Spoken (leader only): set Corruption to 0
\end{itemize}

\subsection*{Table A — Minor Corruption (d12)}
\begin{enumerate}
    \item Cold Breath: Your breath steams in warm rooms. The first time you draw a weapon each scene, tick Alert +1 segment.
    \item Ink-Fingers: Black ink stains your hands. Evidence-type actions suffer DV +1 unless you start by admitting a relevant fault.
    \item Wrong-Whisper: Echoes misname you. Sway vs any harmed party is -1d unless you begin with acknowledgment; on acknowledgment, steady Position.
    \item Guttering Candle: Nearby flames flutter toward you. When you lie about the past, immediately draw the Deck of Consequences.
    \item Numb Grip: Fingers won't warm. Your first delicate action each scene (Tinker/Theft/etc.) starts Risky unless you carry a token of the harmed.
    \item Cold Ring: A band bites your finger. If you break a promise this scene, tick a Memory Collapse [4] sub-clock +1.
    \item Skittering Shadow: Your shadow jitters. Stealth starts Risky unless you spoke a true name today.
    \item Frosted Mirror: Your reflection fogs first. When acting from self-deception, DV +1; if you speak your true name, ignore this for the scene.
    \item Ash in Pocket: Coin spent for selfish gain becomes an IOU; tick Obligation +1 segment (to anyone harmed by the scene).
    \item Blurred Margins: Names blur at the edges of pages. Once/scene the GM may ask, "Whose name is missing?" If you cannot answer, Position -1.
    \item Hollow Step: Floors creak beneath you. Endure-type resists cost +1 stress/strain unless defending the powerless.
    \item Bleeding Ledger: When you destroy a record, mark Stain and roll Minor Corruption again.
\end{enumerate}

\subsection*{Table B — Major Corruption (d12)}
\begin{enumerate}
    \item Counting Shadow: Your shadow counts softly. At Dusk, accept a Compel to attempt a nonviolent option first or mark Stain.
    \item The Unlit Candle: A thirty-first candle won't catch. Undead prefer you as a target, but you gain +1 effect on Rites after confession.
    \item Tongue of Ash: Mispronouncing a true name adds +1 segment to Persuasion/Memory Collapse and locks that name until restitution.
    \item Name-Scar: Healed wounds etch a name you can't say. Choose an NPC's name you are unable to speak until you perform a fitting repair.
    \item Frosted Vows: Your touch chills vows. Allies start the next social scene Risky unless you consecrate with Rune: Witness.
    \item Mirror-Shard Soul: On being reduced to 0, you split into two conflicting drives for one scene (GM tool) until healed by a Vigil.
    \item Debtor's Sigil: Mark Obligation [4] to the Patron. Until cleared, you cannot invoke the same Litany or Glaive name twice per arc.
    \item Name-Eater: On a critical failure, one random name moves from Helped → Unresolved; gain +1d to your immediate follow-up and mark Stain.
    \item Ash Crown: Authority senses your usurpations: DV +1 vs authority until you abdicate something tangible.
    \item Counting Bells: Time slips. Each long action you take also ticks Alert/Ward +1 unless a harmed witness is present.
    \item Empty Pall: Your healing helps half unless accompanied by a spoken apology tied to the scene.
    \item Candle Tax: SB spent against you in a memory-scene has +1 potency (or GM pays 1 less SB) unless a PC reads a harmed name aloud at scene start.
\end{enumerate}

\subsection*{Party Miasma (Optional Group Track)}

When multiple PCs commit erasures in one episode, tick Miasma [0–6].
\begin{itemize}
    \item 3+: All scenes start Position -1 until the group performs a public act of repair together.
    \item 6: The dungeon rejects the party—advance each active clock +1 and introduce a roaming Haunt until you restore a lineage or bury the unburied.
\end{itemize}

\subsection*{Deck Integration (Quick)}

When a result calls for a Deck draw, interpret: Hearts—remembrance, Clubs—pressures, Diamonds—obligations, Spades—omens/necromancy. Use rank for severity.

\subsection*{Runekeeper \& Glaive Interactions}
\begin{itemize}
    \item \textbf{Runekeeper:} Rune: True Name and Rune: Litany cannot be used to harm without offering terms; doing so adds +1 Corruption (then roll Minor).
    \item \textbf{Glaive of the Thirty:} For each point of Corruption, remove one distinct name from being used with Litany Edge until the next arc or until cleansed.
\end{itemize}

\subsection*{GM Usage Notes}
\begin{itemize}
    \item Point to visible marks; make Corruption felt in mirrors, candles, ledgers, and vows.
    \item Make cleansing specific and public; reductions should be stories, not toggles.
    \item Keep the table punchy: roll once, narrate hard, show how mercy re-opens safe play.
\end{itemize}

\end{document}
```