\documentclass[11pt,twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{array}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{framed}
\usepackage{needspace}
\usepackage[table]{xcolor}
\usepackage{hyperref}

% Colors
\definecolor{headerblue}{RGB}{25, 30, 120}
\definecolor{sidebarblue}{RGB}{40, 50, 150}
\definecolor{lightgray}{RGB}{240, 240, 240}
\definecolor{darkgray}{RGB}{80, 80, 80}
\definecolor{tablegray}{gray}{0.9}

% Headers and Footers
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{þepage}
\fancyhead[LO]{Psionics − Fate's Edge Extension}
\fancyhead[RE]{Psionics − Fate's Edge Extension}
\renewcommand{ℎeadrulewidth}{0.4pt}

% Section formatting
\titleformat{\section}
  {≠edspace{4\baselineskip}Łarge\bfseries\color{headerblue}}
  {þesection}{1em}{}
  
\titleformat{\subsection}
  {≠edspace{3\baselineskip}łarge\bfseries\color{sidebarblue}}
  {þesubsection}{1em}{}

% Document info
ℎypersetup{
    colorlinks=true,
    linkcolor=sidebarblue,
    filecolor=magenta,      
    urlcolor=sidebarblue,
}

\title{\textcolor{headerblue}{\Huge Psionics} \\ łarge Extension Module for Fate's Edge}
\author{}
\date{}

\begin{document}

\maketitle
þispagestyle{fancy}

\begin{center}
\begin{minipage}{0.9\textwidth}
\begin{framed}
\noindent \textbf{Design Philosophy:} Psionics in Fate's Edge represents the disciplined mastery of the mind's hidden potentials. Unlike the chaotic forces of magic or the structured pacts of Patrons, psionic power comes from within, shaped by will and honed by practice. This internal source makes psions both powerful and vulnerable−−−their greatest strength is also their most dangerous weakness.
\end{framed}
\end{minipage}
\end{center}

\section*{Psionics: The Mind Ascendant}

In the shadowed courts and ancient ruins where Fate's Edge adventurers tread, power comes in many forms. Some bargain with Patrons, others weave elemental forces, and still others summon spirits from beyond. But there exists a subtler path−−−one that requires no oaths, no incantations, no material components. This is the path of Psionics, where the disciplined mind becomes the greatest tool of all.

Psionics represents the evolution of human potential, the unlocking of abilities that science claims impossible and magic struggles to explain. Where a warlock might call lightning from the sky, a psion shapes reality through the sheer force of thought. Where a mage might scry through mystical mirrors, a psion reaches across space with their consciousness alone.

But this power comes at a cost. The mind that bends reality to its will also bears the scars of that bending. Mental Strain becomes as real a threat as physical exhaustion, and the constant strain of psychic activity can fracture even the strongest psyche.

Based on the playtest experience, here are the key revisions I'd recommend for the Psionics module:

\section{Core Mechanics}

\subsection{Mental Strain Track}
Your \emph{Mental Strain} represents capacity for sustained psychic exertion.

\begin{center}
\textbf{Mental Strain $=$ Spirit}   (minimum 2)
\end{center}

Mark and clear Mental Strain as effects specify. When you \emph{cannot} pay a Mental Strain cost, choose:
\begin{itemize}[leftmargin=*]
  ℹtem Convert the unpaid point to \textbf{+2 Fatigue} (standard system), or
  ℹtem Convert the unpaid point to \textbf{+1 Harm (Stress)}.
\end{itemize}
\textbf{Overflow:} If you ever exceed your Mental Strain, immediately suffer \textbf{+1 Harm (Stress)} and set Mental Strain to your maximum (no further over−marking this scene).

\subsection{Psionics Skill}
\textbf{Psionics (0−−5)} gates access to psionic Arts and sets control/efficiency. You cannot use psionic Arts without investing in this skill.

\subsection{Resolution and Attribute Keys}
Psionics uses \textbf{Attribute (per Art) $+$ Psionics vs.\ DV}. To encourage specialisation, each Art has a \emph{Key Attribute}. Using a non−key attribute for that Art imposes \textbf{+1 DV} (or $−1$ die, GM's choice).

\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{longtable}[c]{@{}lll@{}}
⊤rule
\textbf{Art} & \textbf{Key Attribute} & \textbf{Alternate (with +1 DV)} \\
∣rule
Telekinesis & Spirit (force) & Wits (finesse/precision) \\
Telepathy & Wits & Presence (emotive broadcast) \\
Clairvoyance & Wits & Spirit (ritual trance) \\
Biofeedback & Spirit & Wits (clinical self−hack) \\
Astral Projection & Spirit & Wits (tethered skim) \\
Psychic Assault & Spirit (overwhelm) & Wits (needle, bypass) \\
Mind Shield & Wits (filter) & Spirit (soak) \\
Empathic Manipulation & Presence & Wits (cold read leverage) \\
Precognition & Spirit & Wits (tactical glimpses) \\
⊥tomrule
\end{longtable}
\end{center}

DV is set by the GM (typical 2−−5+) from scope, risk, and opposition.

\section{Psionic Consequence System}

Psionics interfaces with existing \textbf{Fatigue/Harm}. Use \emph{Mental Strain} as the psion−only resource; use \emph{Fatigue/Harm} for spillover consequences.

\subsection{Psionic Strain (simple thresholds)}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Mark Fatigue when:} DV~4+ effects, engaging 3+ foes with psionics in one exchange, or maintaining an effect across 2+ scenes.
  ℹtem \textbf{At 2 Fatigue:} $−1$ die to psionic rolls.
  ℹtem \textbf{At 4 Fatigue:} choose one:
  \begin{itemize}
    ℹtem \textbf{Surge:} One psionic use this scene ignores \emph{Mental Strain} cost; then mark \textbf{+1 Harm} and generate \textbf{+1 SB}.
    ℹtem \textbf{Efficiency:} Halve all \emph{Mental Strain} costs for the next scene (round up).
    ℹtem \textbf{Reset:} Clear all Fatigue; you cannot use psionics for one scene and mark \textbf{+1 Harm}.
  \end{itemize}
\end{itemize}

\section{Psionic Enhancements}
\begin{description}
  ℹtem[Psionic Surge (1/scene):] Spend \textbf{3 Mental Strain} to gain \textbf{+1 Effect} on a psionic action.
  ℹtem[Mental Breakthrough:] At \textbf{4+ Fatigue}, you may Push a psionic effect \textbf{+1 step}; mark \textbf{+2 Harm} and generate \textbf{+2 SB}.
  ℹtem[Psionic Meditation:] During any rest, recover \textbf{+1 Mental Strain} (in addition to normal recovery).
\end{description}

\section{Psionic Arts}

\subsection{Telekinesis}
\textbf{Key Attribute:} Spirit $+$ Psionics  
\textbf{Alternate:} Wits $+$ Psionics (\textbf{+1 DV} for fine manipulation). \\
\textbf{Action:} 1 action; sustain by fiction.  
\textbf{Range:} \textit{Near} (add \textbf{+1 DV} for \textit{Far}). \\
\textbf{Cost:} Activate \textbf{1 Mental Strain}; sustained lifts/holds \textbf{1 Mental Strain/scene}. \\
\textbf{SB:} Scales with mass/force:
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Light (DV 2−−3):} cups, tools, doors. \emph{No SB.}
  ℹtem \textbf{Medium (DV 4):} furniture, a person's mass, riding debris. \emph{+1 SB on Partial/Miss.}
  ℹtem \textbf{Heavy (DV 5+):} carts, statues, boulders, multi−target shoves. \emph{+2 SB on Partial/Miss; +1 SB on Success.}
\end{itemize}
\textbf{Mechanical Benefits:}
\begin{itemize}[leftmargin=*]
  ℹtem Hurl objects as weapons (Harm by fiction and force tier).
  ℹtem Shape cover: lift slabs, hold doors, create transient \emph{Walls/Barriers}.
  ℹtem Remote tasks: pull levers, pick up keys, disarm simple traps at range.
\end{itemize}
\textbf{Limits/Notes:}
\begin{itemize}[leftmargin=*]
  ℹtem Cannot \emph{directly} crush or twist living tissue; injury is \emph{environmental} (falling, debris, impact) at GM discretion.
  ℹtem Complex, delicate work (lockpicks, surgery) typically requires the \textbf{Alternate} attribute (Wits) and adds \textbf{+1 DV}.
\end{itemize}
\textbf{Examples:} Nudge a goblet off a shelf (DV~2); rip a door from hinges (DV~4); wrench a stone pillar sideways (DV~5).

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Telepathy}
\textbf{Key Attribute:} Wits $+$ Psionics  
\textbf{Alternate:} Presence $+$ Psionics (\textbf{+1 DV}, emotive broadcast style). \\
\textbf{Action:} 1 action; sustain by fiction.  
\textbf{Range/Line:} \textit{Near}, LoS (\textbf{+1 DV} for \textit{Far}; \textbf{+1 DV} without LoS if you have a strong link). \\
\textbf{Targets:} One (add \textbf{+1 DV} per extra target this scene). \\
\textbf{Design Rule:} Telepathy never replaces social skills−−−it provides \emph{information and set−up}, not compliance.

\paragraph*{Depths & SB}
Choose a depth before rolling:
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Surface Thoughts (DV 2):} current focus. \emph{No SB} on any result \emph{the first time per target/scene}. Repeats: \textbf{+1 DV}; on Miss, \textbf{+1 SB}.
  ℹtem \textbf{Emotional State (DV 3):} tones/urges. \emph{+1 SB} on Partial/Miss.
  ℹtem \textbf{Deep Memories (DV 5):} episodic recall. \emph{+1 SB} on Success; \emph{+2 SB} on Partial/Miss.
\end{itemize}

\paragraph*{Outcomes}
\begin{description}
  ℹtem[Success:] Gain the info; pick \emph{one} set−up and use it by the end of your next turn this scene:
  \begin{itemize}
    ℹtem \textbf{Angle of Approach:} \( +1\) die to your next \emph{Insight/Sense Motive} vs.\ this target.
    ℹtem \textbf{Read the Room:} \textbf{Position +1} on your next social roll vs.\ this target.
    ℹtem \textbf{Pointed Question:} Ask one focused question they could presently answer; GM replies as impressions.
  \end{itemize}
  ℹtem[Partial:] Blurred fragments; mark SB for depth.
  ℹtem[Miss:] No useful read; mark SB for depth (if any); the target feels the brush.
\end{description}

\paragraph*{Defense & Counterplay}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Unwilling Resist:} target rolls Spirit $+$ Resolve; on meet/beat, your read fails and you mark \textbf{+1 SB} (\textbf{+2 SB} if Deep). Margin \(2+\): they also intuit your intent.
  ℹtem \textbf{Wards/Noise:} \texttt{[WARD]}s, mental static, or anti−psionic training impose \textbf{+1−−2 DV} (training at least \textbf{+1 DV}).
  ℹtem \textbf{Consent:} willing targets are \textbf{DV−−1} and cause no SB.
\end{itemize}

\textbf{Non−Substituting Benefits:}
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Silent Speech (willing only):} short thoughts in \textit{Near} (group size up to your \textbf{Wits}); no SB.
  ℹtem \textbf{Sense Dissonance:} if you read \emph{Surface} this scene, \( +1\) die on \emph{Wits + Insight} to vet a lie; strong hit reveals \emph{which part} conflicts.
\end{itemize}
\textbf{Notes:} Reading unwilling minds in a formal social scene risks fallout; if detected, \textbf{+1 SB} even on Success.

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Clairvoyance}
\textbf{Key Attribute:} Wits $+$ Psionics  
\textbf{Alternate:} Spirit $+$ Psionics (\textbf{+1 DV}, trance method). \\
\textbf{Action:} 1 action; ritual scry may take a scene.  
\textbf{Range:} Scene/locale by fiction; add \textbf{+1 DV} per range band beyond \textit{Near}. \\
\textbf{Cost:} DV 2−−3 = \textbf{1 Mental Strain}; DV 4 = \textbf{2}; DV 5+ = \textbf{3}. \\
\textbf{SB:} \textit{Near} none • \textit{Distant} \textbf{+1 SB} on Partial/Miss • \textit{Complex Scry} \textbf{+2 SB} on Partial/Miss; \textbf{+1 SB} on Success.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Uses:} scout ahead, pierce \texttt{[VEIL]}, track signatures, read recent traces.
  ℹtem \textbf{Notes:} Covert spying on private moments generates \textbf{+1 SB}. Each extra scene sustained adds \textbf{+1 SB}.
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Biofeedback}
\textbf{Key Attribute:} Spirit $+$ Psionics  
\textbf{Alternate:} Wits $+$ Psionics (\textbf{+1 DV}, clinical technique). \\
\textbf{Action:} 1 action (Major work takes a focused scene). \\
\textbf{SB:} by intensity:
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Minor (DV 2):} stabilize, close, purge shock. \emph{No SB.} Heal up to \textbf{Psionics} levels/day (cap 1 Harm per recipient/day).
  ℹtem \textbf{Moderate (DV 3):} \emph{+1 SB} on Partial/Miss; heal up to \textbf{Psionics} (max 2/day).
  ℹtem \textbf{Major (DV 4+):} \emph{+2 SB} on Partial/Miss; \emph{+1 SB} on Success; heal up to \textbf{Psionics} (max 1/day); mark \textbf{+1 Fatigue}.
\end{itemize}
\textbf{Perks:}
\begin{itemize}[leftmargin=*]
  ℹtem Convert \textbf{1 Harm} $↔$ \textbf{1 Mental Strain} (up to \textbf{Psionics}, max 3/day).
  ℹtem Brief enhancement: \textbf{+1 die} to physical rolls for a few moments.
\end{itemize}
\textbf{Limits/Notes:}
\begin{itemize}[leftmargin=*]
  ℹtem Cannot heal Harm above your \textbf{Psionics} rating; active combat imposes \textbf{Position −−1}.
  ℹtem Major healing requires a quiet scene.
  ℹtem Healing or enhancing \emph{others} requires consent and generates \textbf{+1 SB}.
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Astral Projection}
\textbf{Key Attribute:} Spirit $+$ Psionics  
\textbf{Alternate:} Wits $+$ Psionics (\textbf{+1 DV}, skim tether). \\
\textbf{Action:} 1 action to slip; travel by fiction.  
\textbf{Cost:} \textbf{1 Mental Strain/scene}. \\
\textbf{SB:} \textbf{+1} per band beyond \textit{Near}; \textbf{+1} per additional scene sustained.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Benefits:} pass barriers, scout hazards, converse across distance.
  ℹtem \textbf{Vulnerabilities:} body is \textbf{Helpless}; non−physical contests accrue SB; astral Harm echoes to flesh; wards add \textbf{+2 DV}.
  ℹtem \textbf{Risk:} extended ($>3$ scenes) projection courts separation (GM consequence); astral death = physical death.
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Psychic Assault}
\textbf{Key Attribute:} Spirit $+$ Psionics  
\textbf{Alternate:} Wits $+$ Psionics (\textbf{+1 DV}, needle−strike). \\
\textbf{Action:} 1 action.  
\textbf{Cost:} DV 2−−3 = \textbf{1 Mental Strain}; DV 4 = \textbf{2}; DV 5+ = \textbf{3}. \\
\textbf{SB:} always \textbf{+1} (disruptive) plus DV 4 \textbf{+1} and DV 5+ \textbf{+2} more.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Benefits:} mental Harm; bypasses physical armor; target specific faculties (sight, balance, focus) by fiction.
  ℹtem \textbf{Notes:} cannot be "willingly" received; high−Spirit foes may add \textbf{+1 DV}; use in sensitive locales may carry legal or social fallout.
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Mind Shield}
\textbf{Key Attribute:} Wits $+$ Psionics  
\textbf{Alternate:} Spirit $+$ Psionics (\textbf{+1 DV}, soak−and−ground). \\
\textbf{Action:} 1 action (passive modes by fiction). \\
\textbf{SB:} Passive none; \emph{per effect blocked} \textbf{+1 SB}; \emph{Area} \textbf{+2 SB} and mark \textbf{+2 Fatigue}.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Benefits:} resist intrusion/telepathy; shelter allies; create dampened zones.
  ℹtem \textbf{Paradox:} protecting \emph{others} costs \textbf{2$×$ Mental Strain}.
  ℹtem \textbf{Notes:} \emph{Area} protection covers all within \textit{Near} (friends and foes); cannot be maintained while unconscious.
\end{itemize}

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Empathic Manipulation}
\textbf{Key Attribute:} Presence $+$ Psionics  
\textbf{Alternate:} Wits $+$ Psionics (\textbf{+1 DV}, cold−read leverage). \\
\textbf{Action:} 1 action; sustain by fiction. \\
\textbf{SB:} by intensity:
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Subtle (DV 2−−3):} nudge mood, steady hands. \emph{No SB.}
  ℹtem \textbf{Moderate (DV 4):} calm rage, spark courage. \emph{+1 SB on Partial/Miss.}
  ℹtem \textbf{Strong (DV 5+):} sway a crowd's tenor. \emph{+2 SB on Partial/Miss; +1 SB on Success.}
\end{itemize}
\textbf{Benefits:} improve social \textbf{Position +1}; pacify beasts; inspire/demoralize groups. \\
\textbf{Risks/Notes:} You \emph{feel} what you project; same target not more than once per scene; attempts that oppose core values are \textbf{+2 DV}; affecting a crowd always generates \textbf{+1 SB}.

% −−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

\subsection{Precognition}
\textbf{Key Attribute:} Spirit $+$ Psionics  
\textbf{Alternate:} Wits $+$ Psionics (\textbf{+1 DV}, tactical glimpse). \\
\textbf{Action:} 1 action; deeper visions may take a scene. \\
\textbf{Cost:} DV 2−−3 = \textbf{1 Mental Strain}; DV 4 = \textbf{2}; DV 5+ = \textbf{3}. \\
\textbf{SB (inherent):} Minor \textbf{+1} • Major (DV 4) \textbf{+2} • Detailed (DV 5+) \textbf{+3}.
\begin{itemize}[leftmargin=*]
  ℹtem \textbf{Benefits:} advantage on future rolls (scope by GM), avoid named dangers, compare probable branches.
  ℹtem \textbf{Paradox −−− Future Fixation:} until Downtime (or a defiant choice), lose \textbf{1 Boon/session}; actions that resist the shown path may be \textbf{Desperate} by GM call.
  ℹtem \textbf{Notes:} No "pure randomness" predictions (e.g., dice, lots) without \textbf{+3 DV}; visions of death/calamity add \textbf{+1 SB}.
\end{itemize}

\section*{Psionic Costs}

\subsection{Base Costs}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Activation:} 1 Mental Strain
    ℹtem \textbf{Maintenance:} 1 Mental Strain per Scene/Significant Time
    ℹtem \textbf{Skill Requirement:} Must invest XP in Psionics skill
\end{itemize}

\textbf{Revised Note:} Maintenance costs can be reduced by 1 through successful Wits + Psionics (DV 3) test, but failure generates 1 SB.

\section*{Psionic Talents}

\subsection{Minor Talents (2−4 XP)}

\textbf{Mental Fortress (3 XP)} \\
Requirements: Psionics 2+, Spirit 3+ \\
Effect: +1 die to resist psychic effects; once per scene, convert incoming psychic Harm to Fatigue

\textbf{Thought Thief (4 XP)} \\
Requirements: Psionics 2+, Wits 3+, Telepathy \\
Effect: When reading surface thoughts, gain +1 die; once per scene, detect lies automatically

\textbf{Psychic Reservoir (4 XP)} \\
Requirements: Psionics 2+, Spirit 2+ \\
Effect: Increase Mental Strain track by +2; recover 1 additional Mental Strain during rest

\textbf{Revised Note:} All Minor Talents now have a prerequisite of at least 1 other psionic talent, encouraging specialization.

\subsection{Major Talents (6−8 XP)}

\textbf{Telekinetic Mastery (7 XP)} \\
Requirements: Psionics 3+, Spirit 3+, Telekinesis \\
Effect: +1 die to Telekinesis attacks; can lift objects one category heavier without increasing DV

\textbf{Mind Walker (8 XP)} \\
Requirements: Psionics 3+, Wits 3+, Astral Projection \\
Effect: Astral Projection requires no concentration; move at normal speed in astral form; reduce projection cost to 1/2 Mental Strain (round up)

\textbf{Psychic Vampire (8 XP)} \\
Requirements: Psionics 3+, Spirit 3+, Psychic Assault \\
Effect: When dealing psychic Harm, heal 1 Mental Strain; once per scene, drain 1 Mental Strain from target (requires successful Psychic Assault)

\subsection{Prestige Talents (12+ XP)}

\textbf{Master of Minds (12 XP)} \\
Requirements: Psionics 4+, Wits 4+, Telepathy, Empathic Manipulation \\
Effect: +2 dice to all telepathic effects; can communicate with entire group simultaneously; once per scene, impose emotional state on group (DV 4, mark 2 Fatigue)

\textbf{Mind Over Body (12 XP)} \\
Requirements: Psionics 4+, Spirit 4+, Tier III+ \\[2pt]
Effect: Your Mental Strain track becomes \textbf{Spirit $+$ Wits} 
(\emph{min} 6, \emph{cap} 12). Once per scene, when you would suffer Mental Strain \emph{overflow}, you may instead convert each overflow point into \textbf{+1 Fatigue} (standard track), with no per−use limit. \\[2pt]
Paradox: Mind eclipses flesh. When reduced to 0 Harm, you may remain conscious at \textbf{Helpless} until your next turn; then mark \textbf{+2 Fatigue} and \textbf{+1 SB}. \\[2pt]

\subsection*{Reality Bender}
łabel{talent:reality−bender}
∈dex{Talents!Reality Bender}∈dex{Psionics!Reality Bender}

\paragraph*{Type} Major Talent (15 XP)

\paragraph*{Requirements} \textbf{Psionics 5+}, \textbf{Spirit 4+}, \textbf{Telekinesis}, \textbf{Psychic Assault}, \textbf{Mind Shield}.

\paragraph*{Test} \emph{Wits + Psionics} vs.\ \textbf{DV} (per option below).

\paragraph*{Effect} Once per session, reshape local reality within \emph{Near} range. Choose one:
\begin{itemize}
  ℹtem \textbf{Minor Alteration (DV 3):} Change one obvious physical property (color, temperature, texture, sound) of a target within \emph{Close} range, or create a simple illusory effect. \emph{Lasts for the scene.}
  ℹtem \textbf{Moderate Alteration (DV 4):} Alter the structural properties of a \emph{Medium} or smaller object (e.g., make glass flexible, metal brittle, water viscous) for \emph{one exchange}, or create a complex illusion that includes movement and sound.
  ℹtem \textbf{Major Alteration (DV 5):} Temporarily defy a basic law of physics within \emph{Close} range for \emph{one exchange} (e.g., reverse gravity in a small area, stop time for a single object, make fire cold).
\end{itemize}

\paragraph*{Costs} Mark \textbf{1 Harm} and \textbf{1 Fatigue}; generate \textbf{1 SB}.

\paragraph*{Limitations}
\begin{itemize}
  ℹtem Cannot create matter from nothing; you must transform or rephase existing phenomena.
  ℹtem Cannot directly harm creatures (no direct damage; secondary environmental effects at the GM's discretion).
  ℹtem Cannot duplicate other magic paths' signature effects exactly (adjacent expressions at the GM's discretion).
  ℹtem Cannot affect unwilling targets' minds or souls.
\end{itemize}

\paragraph*{Examples} Reverse gravity in a small area; make stone flow like water; create a bridge of force; turn sound to color; make shadows solid.

\textbf{Revised Note:} Reality Bender now generates additional SB based on alteration severity: Minor (+1 SB), Moderate (+2 SB), Major (+3 SB). Use in populated areas generates additional 1 SB.

\section*{Anti−Psionic Effects}

Psionic characters are vulnerable to specific countermeasures that use existing mechanics:

\textbf{Psionic Dampening [WARD]} \\
DV = Psionics skill; Hit: impose −1 die to psionic rolls; Partial: −1 Effect

\textbf{Mental Null Zone} \\
Area effect that prevents psionic use; resisted by Spirit + Psionics

\textbf{Psychic Disruption [BANISH]} \\
Target a psionic effect; DV by fiction; Hit: end effect and mark 1 Fatigue

\textbf{Anti−Psionic Training} \\
Some NPCs get +1 die against psionic effects due to specialized training

\textbf{Revised Note:} Anti−psionic effects are more common in areas with strong magical or technological presence. Professional investigators and security personnel often have basic anti−psionic training.

\section*{Balance Adjustments Summary}

\subsection*{Telepathy}
\begin{itemize}[leftmargin=*]
    ℹtem Limited to once per target per scene for surface thoughts
    ℹtem Increased DV for invasive effects (3−5)
    ℹtem Unwilling targets get +2 DV bonus
    ℹtem Social consequences for obvious use
    ℹtem Legal implications in civilized areas
\end{itemize}

\subsection*{Biofeedback}
\begin{itemize}[leftmargin=*]
    ℹtem Daily limits on healing (1−3 conversions per day based on severity)
    ℹtem Major healing limited to 1 per day and marks Fatigue
    ℹtem Requires concentration for serious healing
    ℹtem Cannot be used freely in combat
    ℹtem Consent required for use on others
\end{itemize}

\subsection*{Psychic Fatigue}
\begin{itemize}[leftmargin=*]
    ℹtem Reset causes 1 Harm
    ℹtem Surge now costs 3 Mental Strain instead of free
    ℹtem Surge generates additional SB
    ℹtem Integrates with existing Fatigue system
\end{itemize}

\subsection*{General}
\begin{itemize}[leftmargin=*]
    ℹtem Integrated with existing Fatigue/Harm system
    ℹtem Uses Story Beats for consequences
    ℹtem Added psionic−specific Talents for advancement
    ℹtem Anti−psionic effects use existing tag system
    ℹtem Maintains parity with other magic paths
    ℹtem Clear talent prerequisite chains
    ℹtem Scaled Fatigue costs for arts
    ℹtem Flexible maintenance costs
    ℹtem Legal and social consequences
\end{itemize}

\section*{Additional Psionic Complications}

\subsection*{Social Integration Issues}
Psionic characters may face prejudice or fear from non−psionic individuals. In some cultures, psionic abilities are viewed with suspicion or regulated by law.

\subsection*{Mental Scars}
Intense psionic use can leave lasting psychological effects. Characters may develop phobias, compulsions, or other mental conditions related to their abilities.

\subsection*{Psionic Signatures}
Powerful psionic use can leave detectable traces that skilled investigators or other psions can track. This creates opportunities for pursuit or identification.

\subsection*{Reality Bleed}
Extensive use of reality−altering abilities may cause minor inconsistencies in the local fabric of reality, creating unpredictable effects that generate Story Beats.

\section*{Psionic Investigation Procedures}

\subsection*{Evidence Handling}
Psionic evidence requires special handling and verification. Mental impressions must be corroborated with physical evidence in most legal systems.

\subsection*{Chain of Custody}
Just as physical evidence requires careful handling, mental evidence must maintain "psychic chain of custody" through proper documentation and witness verification.

\subsection*{Cross−Cultural Considerations}
Different cultures have varying attitudes toward psionic investigation techniques, affecting their admissibility and effectiveness in different jurisdictions.

\section{Cultural Faces of the Silent Art}

≤ttrine{N}{ot} all peoples fear the psion equally, but none trust them fully.

\subsection{Aeler −− The Deep Accountants}
In Aeler, psionics is treated as \textbf{infrastructure}−−−a utility to be maintained, measured, and contained. They assign psions to the Deep Vaults, where their ``breath−debts'' can be monitored. Each Mental Strain point is counted in \textit{bell−rings}−−−a psion carries a bell on a leather cord, and when it rings three times too fast, a Corrector arrives to audit the ledger.

They practice \textbf{Breath Accounting}, converting mental strain into measured respiration. Their psions specialize in \textit{Structural Sense}−−−feeling the stress points in stone−−−and \textit{Resonance Mapping}, sensing the health of the Underways through mental connection.

\subsection{Lethai −− The Context Keepers}
For the Lethai, telepathy is not merely reading thoughts; it requires \textbf{context keys}−−−the proper season, relationship, and place. Without these, a Lethai psion reads only static. Their minds are saturated with memory and social webbing, making them difficult to read (psions who try often leave mad), but also requiring them to navigate complex ritual before mental contact.

They mark their psions with \textbf{Mind−Gifts}−−−resinous tattoos that serve as mental anchors. The Spider's Patience grants concentration; the Serpent's Breath detects deception. A Lethai psion is less a spy and more a \textit{living archive}, but take them from their context and they are deafened.

\subsection{Ykrul −− The Weather Readers}
The orcs of the steppe do not see the future in tea leaves. They see it in the \textbf{Kon'reh}−−−the sacred geometry of probability. Their psions are not called seers but \textit{Riders of the Flow}. They read the shape of the future as a cavalryman reads the shape of the land: tactical, immediate, and bound to the weather.

Storm conditions enhance their telepathic range but increase Mental Strain costs. Clear weather allows precise precognitive readings. They specialize in \textit{Herd Sense}−−−feeling the emotional state of a clan−−−and \textit{Banner Communication}, long−distance mental coordination between scattered riders.

\subsection{Vilakari −− The Information Brokers}
In the free ports, psionics is commerce. The Vilakari have developed \textbf{Thought Markets}, where memory and information are traded like grain. Their psions are \textit{Surface Readers}, adept at detecting deception in mercantile negotiations, and \textit{Memory Traders}, who temporarily share specific recollections for coin.

But their gift is also their vulnerability. A Vilakari psion in a crowded market is overwhelmed by the noise of surface thoughts. They require specialized training to filter the signal from the cacophony, and many fall to addiction−−−craving the rush of other minds, burning out their own Mental Strain in weeks rather than years.

% −−− NEW SECTION: SILENT ORDERS (replaces the old brief version) −−−

\section{The Silent Orders −−− Fragments of a Scattered Tradition}

\noindent No one knows where psionics came from. The Thepyrgos Synod has a standing bounty for any evidence dated to the Great Silence. The bounty has never been claimed. What follows are whispers collected by the Grey Wanderer from a dozen dusty archives and one very expensive bottle of brandy.

\subsection*{The Scholarly Debate −−− As Reported by the Grey Wanderer}

\begin{quote}
\emph{"The Silence is not an event. It is a lacuna. A gap in the record that has been filled with too many theories and not enough evidence."} \\
−−− Keeper Lian, Lethai−thora, from a lecture at the University of Thepyrgos
\end{quote}

The scholars argue. They always argue.

\begin{description}
  ℹtem[The Withdrawal (Lethai−thora tradition).] Some say the ancient psions did not fall. They \textit{chose} to be silent. They learned that consciousness without discipline attracts something from the spaces between thoughts. The Silence was a quarantine, not a catastrophe. (The Aeler dismiss this as "elves romanticizing their own amnesia.")

  ℹtem[The Broken Frequency (Aeler vent−priors).] Others point to resonance. You cannot "choose" to stop breathing, they say. Something struck the harmonic balance of the mind. A weapon? A natural collapse? The deep vaults still ring with a frequency that does not decay. The Silence is not silence. It is \textit{interference}. (The Lethai−thora note that the Aeler have never produced a single shard of this supposed "weapon.")

  ℹtem[The Forgotten Patron (anonymous Invoker of The Witness).] A third camp whispers of a patron whose domain is \textit{forgetting}. Not Ikasha−−−she sleeps. Not the Ninth−−−it omits. This one \textit{erases}. The ancient psions got too close to its true name. The Silence is the sound of that name being spoken once, then never again. (The Synod has classified this theory as "speculative and possibly heretical.")

  ℹtem[The Empty Circle (Kuvani oral tradition, transcribed by an Aelinnel scribe).] The wanderers of the eastern steppe say: "They did not die. They walked into the Empty Circle and kept walking. The Circle is not a hole. It is a \textit{door}. They are still walking. When they return, the Silence will end." (The Empty Circle is real. It nullifies all psionics. No one has ever walked out of its center.)

  ℹtem[The Simple Answer (skeptical Thepyrgos clerk).] Perhaps they simply exhausted themselves. They had no patrons, no codices, no symbols. Their power was purely internal. Internal things fail from internal causes. They forgot how to be what they were. (The Grey Wanderer notes that this explanation is the least satisfying and therefore the most likely to be true.)
\end{description}

No one knows. The Silence remains silent.

\subsection*{The Orders That Remain}

Whatever caused the Silence, it did not destroy everything. Scattered fragments survive−−−not as an empire, but as small, secretive \textbf{orders} of monks who keep old disciplines alive. They are not wizards. They are \textit{meditants}. They do not bargain with Patrons. They bargain with themselves.

Three such orders are known to the Grey Wanderer. There may be more.

⫕section{The Order of the Unstruck Bell (Aelerian Deep Vaults)}

These monks dwell in sealed galleries beneath the oldest Aeler holds. The Aeler tolerate them because they maintain "quiet zones" where stone−press is less aggressive. They use \textit{null−bells}−−−bronze bells that ring only in the mind.

\textbf{Player Benefits (Unstruck Bell Monk):}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Null−Bell Bond:} You carry a small bronze bell that only you can hear. It counts as a \textit{String} (once per session, ring it to gain +1 Position on a single Mind Shield or Telepathy roll).
    ℹtem \textbf{Resonant Exhale (Technique):} Once per scene, reduce your Mental Strain by 1 (to a minimum of 0). The GM immediately gains 1 Story Beat, which they may spend on a complication tied to sound, vibration, or a distant chime.
    ℹtem \textbf{Gallery−Bound:} You are comfortable in enclosed stone spaces. You ignore the first level of environmental penalty from underground or confined areas when using psionic Arts.
\end{itemize}

⫕section{The Order of the Empty Circle (Violet Steppe)}

These monks live in tents around the Empty Circle−−−a depression in the steppe where no psionics work. They enter the Circle only to meditate in absolute silence. They believe the Silence was a \textit{lesson}, not a wound.

\textbf{Player Benefits (Empty Circle Monk):}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Circle−Trained:} Once per session, when you would generate a Story Beat from Mental Strain on a non−psionic roll, you may cancel that SB by marking 1 Fatigue (the strain of enforced stillness).
    ℹtem \textbf{Still Mind (Technique):} For the duration of a scene, you may choose to operate at a reduced capacity. While active, any non−psionic action you take generates 1 fewer SB from Mental Strain (minimum 0), but you cannot use any psionic Art with a Tier higher than I (i.e., only minor effects, DV 2−−3). This technique requires a free action to activate and deactivate.
    ℹtem \textbf{Herd Sense:} You gain +1 die to detect emotional states (anger, fear, calm) in any creature within Near range. This is a passive ability.
\end{itemize}

⫕section{The Order of the Shared Breath (Eastern Isles)}

These monks practice \textit{paired meditation}−−−two or more psions sharing Mental Strain across a Circle. They believe that debt is not individual but \textit{communal}.

\textbf{Player Benefits (Shared Breath Monk):}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Circle Bond:} You may form a mental link (the "Circle") with up to three willing allies (they need not be psions, but only psions can share Mental Strain). The link lasts for one scene and requires a free action to initiate.
    ℹtem \textbf{Shared Burden (Technique):} When you generate Mental Strain, you may transfer any amount of it (up to your Spirit) to another member of your Circle. That member must be within Near range and willing. The GM gains 1 Story Beat each time you use this technique.
    ℹtem \textbf{Circle's Breath:} While linked, you and your Circle members gain +1 die to all psionic rolls, but if one member fails a Resolve test to resist fear or mental intrusion, all members test at +1 DV.
\end{itemize}

\noindent \textit{Note:} These orders are not factions in the usual sense. They have no grand agenda. They simply survive, train, and occasionally produce a wanderer who walks the roads with nothing but a bell, a stone, or a breath.

\vspace{0.5cm}
ℎfill −−− \textit{The Grey Wanderer, from a conversation at the Dusty Page}

% −−− END OF NEW SECTION −−−

\section{The Thought Trade: Psionic Economics}

In the bustling markets of Zakov's free ports and the ancient bazaars of Silkstrand, a new form of commerce has emerged that challenges traditional notions of value and exchange. The Thought Trade represents the economic system that has grown around psionic abilities, where mental services, experiences, and even raw consciousness itself have become commodities with measurable worth.

\subsection{Mental Labor Markets}

Just as physical labor has wages and specialized skills, psionic abilities have created entirely new categories of mental labor that command premium prices across all social strata:

⫕section{Memory Merchants}
Telepaths who specialize in memory storage and retrieval have become invaluable to merchants, scholars, and nobles who need to preserve sensitive information without physical records. The going rate for one hour of secure memory storage is equivalent to a skilled craftsman's daily wage, but the risk premium for storing particularly dangerous or controversial memories can increase the price tenfold.

\textbf{Market Integration:} In cities like Ecktoria and Thepyrgos, Memory Guilds have established standardized contracts that specify the exact nature of memories to be stored, the duration of storage, and the conditions under which memories can be retrieved. Breaking a memory storage contract is considered a serious offense, often resulting in both legal penalties and social ostracism that can last for generations.

⫕section{Calculation Circles}
Psions with enhanced mathematical abilities have formed Calculation Circles that serve as living computers for complex engineering, navigation, and financial projects. These psions can perform calculations that would take traditional mathematicians months in a matter of hours, but the mental strain often requires significant recovery time.

\textbf{Economic Impact:} The rise of Calculation Circles has dramatically reduced the time and cost of large−scale projects, from Aeler bridge construction to naval navigation for the Amaranthine Sea fleets. However, it has also created a dependency on psionic labor that some traditional craftsmen view as a threat to their livelihoods.

⫕section{Emotional Architects}
Empathic manipulators who specialize in mood and atmosphere control have found employment in everything from political rallies to theatrical performances. Their ability to subtly influence crowd emotions has made them highly sought after by politicians, merchants, and entertainers.

\textbf{Regulatory Response:} Several city−states have begun implementing "Mood Licensing" requirements for Emotional Architects, particularly when their services are used in political or commercial contexts. Violations can result in both financial penalties and temporary suspension of psionic practice licenses.

\subsection{Thought Currency Systems}

The most revolutionary aspect of the Thought Trade is the development of systems where mental experiences themselves function as currency:

⫕section{Experience Exchange}
In certain underground markets, particularly in Free Havens like Sable Key, psions can trade raw experiences−−−memories of adventure, intense emotions, or unique sensations−−−as direct equivalents to coin. A memory of a dragon flight might be worth fifty silver pieces, while the experience of true love might be valued at several gold coins.

\textbf{Market Dynamics:} The value of experiences fluctuates based on rarity, intensity, and cultural significance. Memories of historical events, first contact with supernatural entities, or particularly intense emotional experiences command premium prices. However, the market is also subject to fraud−−−false or artificially enhanced memories can be difficult to detect and have led to the development of specialized "Memory Authenticators."

⫕section{Skill Debt}
Psions can enter into "Skill Debt" arrangements where they agree to provide specific mental services over time in exchange for immediate payment or favors. This has created a complex system of mental credit and obligation that operates alongside traditional economic systems.

\textbf{Contract Law:} The legal frameworks around Skill Debt vary significantly between jurisdictions. In Aeler holds, such contracts are treated with the same seriousness as infrastructure agreements, with detailed terms and severe penalties for breach. In more chaotic environments like Free Havens, Skill Debt arrangements may be more informal but carry higher risks for both parties.

\subsection{Psionic Guilds and Economic Power}

The economic importance of psionic abilities has led to the formation of powerful guilds that control various aspects of the Thought Trade:

⫕section{The Memory Merchants' Consortium}
Operating primarily in the major port cities, this guild controls the legitimate market for memory storage and retrieval services. They maintain strict quality standards, detailed record−keeping systems, and significant political influence in cities where information security is paramount.

\textbf{Guild Benefits:} Members receive access to secure storage facilities, legal protection for their services, and networking opportunities with high−value clients. However, they must also adhere to strict confidentiality agreements and contribute a percentage of their earnings to guild funds.

⫕section{The Calculation Collective}
A loose network of mathematical psions who coordinate on large−scale projects that require multiple specialists. They've developed standardized protocols for collaborative mental computation and maintain a reputation for accuracy that commands premium rates.

\textbf{Collective Structure:} Unlike traditional guilds, the Calculation Collective operates more like a cooperative, with members sharing both risks and rewards. This structure has proven particularly effective for projects that require sustained mental effort from multiple participants.

\subsection{Economic Integration with Other Modules}

The Thought Trade creates natural connections with other aspects of the Fate's Edge economy:

⫕section{Caravan Integration}
Psionic couriers who can carry information mentally rather than physically have revolutionized long−distance communication. A telepath can carry the equivalent of an entire library across dangerous terrain without risk of physical loss, but the mental strain limits the duration and frequency of such services.

\textbf{Market Impact:} This has created new trade routes and commercial relationships, as merchants can now conduct business across vast distances with unprecedented speed and security. However, it has also made psionic couriers targets for kidnapping and exploitation.

⫕section{Political Intrigue Connections}
The ability to read minds and influence emotions has made psions both valuable assets and dangerous threats in political contexts. Political factions invest heavily in psionic protection (mental shielding) while also seeking to recruit or compromise opposing psions.

\textbf{Diplomatic Protocol:} Many courts have established formal protocols for psionic interactions, including designated "neutral mind spaces" where telepaths can operate without advantage to either side, and strict rules about when and how mental abilities can be used in negotiations.

⫕section{Black Banner Applications}
Military organizations have integrated psionic units into their command structures, creating new tactical possibilities and logistical challenges. Psionic reconnaissance, communication, and psychological warfare capabilities have become essential elements of modern military strategy.

\textbf{Military Economics:} The cost of training and maintaining psionic military units is extremely high, but the strategic advantages often justify the investment. This has created a new category of military expenditure and influenced the balance of power between different factions.

\subsection{Social and Cultural Implications}

The Thought Trade has created new social dynamics and cultural practices:

⫕section{The Gentry of the Mind}
A new social class has emerged consisting of wealthy individuals who have purchased extensive mental experiences, creating what some call "experiential aristocracy." These individuals often have more diverse and intense life experiences than those who rely solely on physical adventures or traditional education.

\textbf{Social Tension:} This has created friction with traditional nobility and wealthy merchants who view mental experiences as somehow less authentic than physical achievements. However, the practical advantages of extensive mental training and experience have made this tension less significant in practical terms.

⫕section{The Privacy Divide}
Society has split between those who have access to mental privacy protection (through anti−psionic training, magical wards, or technological means) and those who do not. This has created new forms of social stratification and economic inequality.

\textbf{Market Response:} Various industries have emerged to serve the privacy protection market, from specialized craftsmen who create anti−telepathic jewelry to legal specialists who navigate the complex regulations around mental privacy rights.

The Thought Trade represents one of the most significant economic innovations in the Fate's Edge setting, demonstrating how psionic abilities can create entirely new categories of value and exchange while also generating new forms of social complexity and economic tension. For players, this system provides numerous opportunities for adventure, from protecting valuable mental cargo to navigating the complex legal and ethical issues that arise when consciousness itself becomes a commodity.

\section{The Great Silence: Psionic History}

In the ancient past, before the rise of the Aeler holds and the establishment of the first Ykrul moots, the world experienced what scholars now call the Great Silence−−−a period when psionic abilities across all peoples were dramatically altered, suppressed, or transformed. This event, which lasted for what contemporary accounts describe as "three turnings of the great wheel" (approximately 200−300 years), fundamentally changed how minds could touch minds across the world.

\subsection{The Silence Event}

The Great Silence did not eliminate psionic abilities entirely, but rather \textit{compressed} them−−−making mental powers more difficult to access, more costly to maintain, and more dangerous to use. During this period, what had once been natural gifts became rare talents that required intense training and came with significant risks.

Archaeological evidence from this period shows:
\begin{itemize}
    ℹtem Abandoned psionic academies across multiple cultures
    ℹtem Sealed mental archives that were too dangerous to access
    ℹtem Artifacts designed to \textit{suppress} rather than enhance mental abilities
    ℹtem Records of "mind−storms" where uncontrolled psionic energy caused physical and social chaos
\end{itemize}

\subsection{Resonance Scars}

Certain locations still bear the marks of the Great Silence, creating what scholars term "Resonance Scars"−−−areas where psionic abilities function differently due to the lingering effects of that ancient suppression:

⫕section{The Whispering Vaults (Aeler Holds)}
Deep within the oldest Aeler underways, certain chambers amplify Mental Strain by 2 segments but also grant +1 Effect to telepathic abilities. These spaces are both coveted and feared by Aeler psions, who must weigh the enhanced abilities against the increased mental cost.

⫕section{The Empty Circle (Violet Steppe)}
A circular depression in the heart of Ykrul territory where no psionic abilities function at all. Ykrul oral history claims this was once the site of the "Last Conclave"−−−where the greatest psions of their ancestors attempted to end the Silence but instead created a permanent null zone.

⫕section{The Singing Stones (Lethai Valewood)}
Ancient standing stones that resonate with psionic energy, allowing for enhanced telepathic communication across vast distances−−−but only when the stones themselves are "tuned" through proper context and ceremony.

\subsection{Awakening Cycles}

Since the end of the Great Silence, psionic abilities have slowly returned in what scholars recognize as distinct \textit{Awakening Cycles}. These periods of increased psionic activity last 50−75 years and are separated by longer periods of relative dormancy:

\begin{description}
    ℹtem[First Awakening (Current Cycle, 150 years ago − present)] Marked by the gradual return of basic telepathic and telekinetic abilities, particularly among younger generations.
    
    ℹtem[Second Awakening (400−325 years ago)] The period when precognition and biofeedback abilities first reappeared, leading to the establishment of many current psionic traditions.
    
    ℹtem[Third Awakening (650−575 years ago)] The mysterious period when astral projection and empathic manipulation returned, though records from this time are fragmentary.
\end{description}

Each Awakening Cycle brings not just the return of abilities, but the transformation of how those abilities function. Modern psions often find that their powers work differently from historical accounts, suggesting that each cycle fundamentally reshapes the nature of psionic interaction.

\subsection{Cultural Memory and the Silence}

Different cultures retain distinct memories of the Great Silence, each reflecting their unique relationship with mental abilities:

⫕section{Aeler Records: The Accounting}
Aeler historical texts describe the Silence as a "misalignment of the mental ledgers"−−−where the balance between mind and world became corrupted. They maintain that the current Awakening Cycle requires careful "mental accounting" to prevent another collapse.

⫕section{Ykrul Oral Tradition: The Broken Paths}
Ykrul stories speak of the Silence as a time when "all routes between minds became blocked," making their Kon'reh−based approach to mental interaction impossible. They believe the current cycle represents a "reweaving" of those mental paths.

⫕section{Lethai Codices: The Lost Context}
Lethai records suggest that during the Silence, the context−saturated nature of their communication was temporarily severed, making their gift−based society impossible. They view the current Awakening as a "recontextualization" of mental abilities.

\subsection{Modern Implications}

The Great Silence has several ongoing effects on psionic practice in the present day:

\begin{enumerate}
    ℹtem \textbf{Generational Differences:} Older psions often have different abilities and limitations compared to those born during the current Awakening Cycle.
    
    ℹtem \textbf{Cultural Tensions:} Different peoples have varying levels of comfort with psionic abilities based on their historical relationship with the Silence.
    
    ℹtem \textbf{Scholarly Interest:} The study of Resonance Scars and Awakening Cycles has become a major field of research, with implications for understanding the fundamental nature of consciousness itself.
    
    ℹtem \textbf{Prophetic Concerns:} Some scholars worry that the current Awakening Cycle may be building toward another Silence event, creating tension between those who want to preserve psionic abilities and those who fear their consequences.
\end{enumerate}

The Great Silence serves as a reminder that psionic abilities are not static−−−they are part of a dynamic system that has shaped and been shaped by the history of the world. For psionic characters, this creates both opportunities and dangers, as they may be part of a new transformation in how minds can touch minds across the Fate's Edge setting.

\begin{center}
\begin{minipage}{0.9\textwidth}
\begin{framed}
\noindent \textbf{Design Philosophy:} This bestiary and generator system provides flexible tools for creating diverse psionic encounters that maintain the unique flavor and mechanical identity of the Fate's Edge psionics system. Each entry emphasizes psychic abilities and consequences that integrate seamlessly with the core module.
\end{framed}
\end{minipage}
\end{center}

\appendix

\section{Psychic Bestiary}

\subsection{Minor Psychic Threats}

⫕section{Mind Tick}
\textbf{Tiny psychic parasite that feeds on stray thoughts}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Street
    ℹtem \textbf{Attributes:} Body 1, Wits 3, Spirit 2
    ℹtem \textbf{Skills:} Stealth 2, Insight 1
    ℹtem \textbf{Psionics:} 1 (Telepathy)
    ℹtem \textbf{Harm:} 1 (Mental Strain drain)
    ℹtem \textbf{Special:} Can phase through small openings, detects psychic activity
    ℹtem \textbf{Tactics:} Sneak attacks on sleeping or meditating targets
\end{itemize}

⫕section{Echo Wraith}
\textbf{Fragment of psychic energy from traumatic events}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Street
    ℹtem \textbf{Attributes:} Body 1, Wits 2, Spirit 4
    ℹtem \textbf{Skills:} Stealth 2, Survival 1
    ℹtem \textbf{Psionics:} 2 (Telepathy, Clairvoyance)
    ℹtem \textbf{Harm:} 2 (psychic assault)
    ℹtem \textbf{Special:} Invisible in areas of psychic trauma, feeds on fear
    ℹtem \textbf{Tactics:} Haunt locations, induce panic to strengthen itself
\end{itemize}

⫕section{Thought Thief}
\textbf{Humanoid who has learned basic mental intrusion}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Street
    ℹtem \textbf{Attributes:} Body 2, Wits 3, Spirit 2
    ℹtem \textbf{Skills:} Sway 2, Deception 2
    ℹtem \textbf{Psionics:} 2 (Telepathy, Empathic Manipulation)
    ℹtem \textbf{Harm:} 1 (physical), 1 (mental)
    ℹtem \textbf{Special:} Can steal surface memories, resistant to mental detection
    ℹtem \textbf{Tactics:} Social manipulation, memory extraction
\end{itemize}

\subsection{Moderate Psychic Threats}

⫕section{Psi−Stalker}
\textbf{Professional hunter who uses telepathy to track targets}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Guild
    ℹtem \textbf{Attributes:} Body 3, Wits 4, Spirit 3
    ℹtem \textbf{Skills:} Stealth 3, Survival 2, Insight 2
    ℹtem \textbf{Psionics:} 3 (Telepathy, Clairvoyance, Precognition)
    ℹtem \textbf{Harm:} 2 (physical), 1 (mental)
    ℹtem \textbf{Special:} Can track psychic signatures across miles, precognitive danger sense
    ℹtem \textbf{Tactics:} Patient hunting, strike when target is vulnerable
\end{itemize}

⫕section{Mentalist Cultist}
\textbf{Devotee of psionic philosophy who channels group power}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Guild
    ℹtem \textbf{Attributes:} Body 2, Wits 3, Spirit 4
    ℹtem \textbf{Skills:} Command 2, Lore 2, Sway 2
    ℹtem \textbf{Psionics:} 3 (Telepathy, Empathic Manipulation, Mind Shield)
    ℹtem \textbf{Harm:} 1 (physical), 2 (psychic assault)
    ℹtem \textbf{Special:} Stronger when near other cultists, can share mental states
    ℹtem \textbf{Tactics:} Crowd control, mental domination, group coordination
\end{itemize}

⫕section{Astral Predator}
\textbf{Creature that exists partially outside normal reality}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Guild
    ℹtem \textbf{Attributes:} Body 2, Wits 3, Spirit 5
    ℹtem \textbf{Skills:} Stealth 3, Athletics 2
    ℹtem \textbf{Psionics:} 3 (Astral Projection, Telekinesis, Mind Shield)
    ℹtem \textbf{Harm:} 2 (physical), 2 (psychic)
    ℹtem \textbf{Special:} Can attack from astral form, immune to non−magical weapons
    ℹtem \textbf{Tactics:} Phase between realms, surprise attacks, ignore physical defenses
\end{itemize}

\subsection{Major Psychic Threats}

⫕section{Corrupted Psion}
\textbf{Former practitioner driven mad by power}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} City
    ℹtem \textbf{Attributes:} Body 3, Wits 4, Spirit 5
    ℹtem \textbf{Skills:} Melee 2, Insight 3, Survival 2
    ℹtem \textbf{Psionics:} 4 (All Arts at reduced effectiveness)
    ℹtem \textbf{Harm:} 2 (physical), 3 (psychic assault)
    ℹtem \textbf{Special:} Psychic Contamination full, causes SB on nearby psions, unstable powers
    ℹtem \textbf{Tactics:} Erratic behavior, overwhelming psychic assault, self−destructive power use
\end{itemize}

⫕section{Mind Archon}
\textbf{Psionic entity that has transcended physical form}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} City
    ℹtem \textbf{Attributes:} Body 1, Wits 5, Spirit 6
    ℹtem \textbf{Skills:} Command 4, Lore 4
    ℹtem \textbf{Psionics:} 5 (Telepathy, Clairvoyance, Psychic Assault, Mind Shield, Precognition)
    ℹtem \textbf{Harm:} 3 (psychic assault), can dominate minds
    ℹtem \textbf{Special:} Exists primarily astrally, controls lesser psychic entities, vast mental reach
    ℹtem \textbf{Tactics:} Mental domination, puppeteering minions, precognitive strategic advantage
\end{itemize}

⫕section{The Hollow Crown}
\textbf{Psionic artifact that has gained malevolent consciousness}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Crown
    ℹtem \textbf{Attributes:} Body 2, Wits 6, Spirit 6
    ℹtem \textbf{Skills:} Command 5, Sway 4
    ℹtem \textbf{Psionics:} 5 (All Arts at full strength)
    ℹtem \textbf{Harm:} 3 (psychic assault), can impose mental conditions
    ℹtem \textbf{Special:} Influences all within psychic range, resistant to physical damage, grows stronger with use
    ℹtem \textbf{Tactics:} Subtle manipulation, turning allies against each other, reality warping
\end{itemize}

\subsection{Legendary Psychic Threats}

⫕section{The Fractured Sage}
\textbf{Ancient psion who has split consciousness across multiple bodies}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Tier:} Synod
    ℹtem \textbf{Attributes:} Body 4 (each fragment), Wits 6, Spirit 6
    ℹtem \textbf{Skills:} All skills 3+
    ℹtem \textbf{Psionics:} 5 (Each fragment specializes in different Arts)
    ℹtem \textbf{Harm:} 3 (physical), 3 (psychic), can regenerate from mental essence
    ℹtem \textbf{Special:} Multiple simultaneous presences, shared consciousness, nearly immortal
    ℹtem \textbf{Tactics:} Divide and conquer, overwhelming coordination, impossible to truly defeat
\end{itemize}

\section{Psionic Encounter Generators}

\subsection{Psychic Threat Generator (d100)}

\begin{center}
\rowcolors{2}{tablegray}{white}
\begin{longtable}{|c|l|p{7cm}|}
ℎline
\rowcolor{headerblue!10}
\textbf{Roll} & \textbf{Threat Type} & \textbf{Description} \\
ℎline
01−10 & Minor Threat & 1 creature, basic psionic abilities \\
11−30 & Minor Group & 2−3 creatures, coordinated tactics \\
31−50 & Moderate Threat & 1 creature, significant psionic abilities \\
51−70 & Moderate Group & Small group with enhanced capabilities \\
71−85 & Major Threat & 1 powerful creature, serious challenge \\
86−95 & Major Complex & Powerful threat with minions/support \\
96−00 & Legendary Threat & Campaign−level psychic entity \\
ℎline
\end{longtable}
\end{center}

\subsection{Psychic Environment Generator (d12)}

\begin{enumerate}[leftmargin=*]
    ℹtem \textbf{Psychic Dead Zone:} No psionic abilities function, Mental Strain cannot be spent
    ℹtem \textbf{Mental Resonance Chamber:} All psionic use marks +1 Mental Resonance
    ℹtem \textbf{Echo Corridor:} Past psionic events replay as illusions
    ℹtem \textbf{Thought Storm:} Random telepathic messages bombard all minds
    ℹtem \textbf{Astral Convergence:} Boundary between realms becomes thin
    ℹtem \textbf{Memory Maze:} Physical space shifts based on memories
    ℹtem \textbf{Emotion Vortex:} Strong feelings become contagious and amplified
    ℹtem \textbf{Precognition Field:} Brief glimpses of immediate future
    ℹtem \textbf{Psychic Feedback Loop:} Psionic effects echo and multiply
    ℹtem \textbf{Mind Mirror:} Thoughts become visible to others
    ℹtem \textbf{Willpower Drain:} All mental actions cost +1 Mental Strain
    ℹtem \textbf{Clairvoyant Nexus:} All hidden things become partially visible
\end{enumerate}

\subsection{Psionic Artifact Generator (d12)}

\begin{enumerate}[leftmargin=*]
    ℹtem \textbf{Thought Recorder:} Captures and replays memories of users
    ℹtem \textbf{Emotion Amplifier:} Intensifies feelings in area
    ℹtem \textbf{Astral Compass:} Points toward nearest psychic signature
    ℹtem \textbf{Memory Crystal:} Stores vast amounts of information psionically
    ℹtem \textbf{Precognition Shard:} Grants brief future visions but marks Echo Traces
    ℹtem \textbf{Mind Shield Generator:} Creates area of psychic protection
    ℹtem \textbf{Telepathy Amplifier:} Extends range and power of mental communication
    ℹtem \textbf{Psychic Battery:} Stores Mental Strain for later use
    ℹtem \textbf{Reality Anchor:} Prevents astral projection and psychic phase−shifting
    ℹtem \textbf{Contamination Core:} Concentrates psychic corruption effects
    ℹtem \textbf{Resonance Tuner:} Allows multiple psions to share Mental Resonance benefits
    ℹtem \textbf{Null Stone:} Temporarily suppresses all psionic abilities in area
\end{enumerate}

\subsection{Psychic Complication Generator (d12)}

\begin{enumerate}[leftmargin=*]
    ℹtem \textbf{Tracking Signature:} Your psychic use leaves detectable traces
    ℹtem \textbf{Mental Parasite:} Foreign consciousness shares your mind
    ℹtem \textbf{Power Instability:} Psionic abilities trigger randomly under stress
    ℹtem \textbf{Empathic Overflow:} You constantly feel others' emotions
    ℹtem \textbf{Temporal Echo:} You occasionally glimpse possible futures
    ℹtem \textbf{Psychic Addiction:} You crave psionic use, suffer penalties without it
    ℹtem \textbf{Mental Scars:} Past trauma affects current psionic control
    ℹtem \textbf{Resonance Sensitivity:} You're overwhelmed by other psions' activities
    ℹtem \textbf{Astral Bleed:} Your physical and astral forms don't align properly
    ℹtem \textbf{Memory Fragmentation:} Psionic use causes gaps in your memories
    ℹtem \textbf{Emotional Contagion:} Your emotions strongly affect those nearby
    ℹtem \textbf{Future Fixation:} Precognition makes you feel trapped in destiny
\end{enumerate}

\subsection{Psionic Patron Generator (d8)}

\begin{enumerate}[leftmargin=*]
    ℹtem \textbf{The Silent Order:} Ancient psionic brotherhood seeking balance
    ℹtem \textbf{The Mind Merchants:} Trade in psychic services and information
    ℹtem \textbf{The Astral Explorers:} Seek knowledge beyond physical reality
    ℹtem \textbf{The Thought Police:} Enforce laws regarding mental privacy
    ℹtem \textbf{The Emotion Artists:} Manipulate feelings as creative expression
    ℹtem \textbf{The Precognition Traders:} Profit from glimpses of the future
    ℹtem \textbf{The Memory Keepers:} Preserve and protect important knowledge
    ℹtem \textbf{The Fractured Collective:} Hive−mind seeking to expand consciousness
\end{enumerate}

\subsection{Psionic Organization Generator (d8)}

\begin{enumerate}[leftmargin=*]
    ℹtem \textbf{The Quiet Schools:} Educational institutions for psionic development
    ℹtem \textbf{The Mind Merchants Guild:} Commercial organization trading psychic services
    ℹtem \textbf{The Silent Hunters:} Professional trackers using telepathy
    ℹtem \textbf{The Astral Guard:} Protectors who fight interdimensional threats
    ℹtem \textbf{The Thought Police:} Law enforcement specializing in mental crimes
    ℹtem \textbf{The Emotion Theaters:} Performance groups using empathic manipulation
    ℹtem \textbf{The Future Traders:} Economic consortium profiting from precognition
    ℹtem \textbf{The Memory Libraries:} Archives preserving knowledge psionically
\end{enumerate}

\section{Psionic Encounter Building}

\subsection{Single Threat Encounter}
\begin{enumerate}[leftmargin=*]
    ℹtem Roll Psychic Threat Generator
    ℹtem Roll Psychic Environment Generator for location modifier
    ℹtem Roll 1d4 for additional complications (max 2)
    ℹtem Roll 1d6 for beneficial environmental effects (1−2 = none, 3−4 = minor, 5−6 = significant)
\end{enumerate}

\subsection{Organization Encounter}
\begin{enumerate}[leftmargin=*]
    ℹtem Roll Psychic Organization Generator
    ℹtem Roll 2d4 for number of members (2−3 = patrol, 4−5 = squad, 6−8 = strike team)
    ℹtem Roll 1d3 for leader type (standard member, enhanced member, psionic specialist)
    ℹtem Roll Psychic Environment Generator for operational theater
    ℹtem Roll 1d4 for organizational goal (patrol, capture, eliminate, retrieve)
\end{enumerate}

\subsection{Artifact Encounter}
\begin{enumerate}[leftmargin=*]
    ℹtem Roll Psionic Artifact Generator
    ℹtem Determine artifact's current state (1−4 = dormant, 5−8 = active, 9−12 = malevolent)
    ℹtem Roll 1d6 for artifact's guardian (1−2 = none, 3−4 = minor threat, 5 = moderate threat, 6 = major threat)
    ℹtem Roll Psychic Environment Generator for artifact's location
    ℹtem Roll 1d4 for artifact's effect on users (1 = beneficial, 2 = neutral, 3 = harmful, 4 = transformative)
\end{enumerate}

\section{Psionic Campaign Threads}

\subsection{The Resonance War}
\textbf{Two psionic organizations are escalating their conflict}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Minor Threats:} 3−5 encounters with organization members
    ℹtem \textbf{Moderate Threats:} 2−3 encounters with specialists
    ℹtem \textbf{Major Threat:} 1 encounter with organization leadership
    ℹtem \textbf{Campaign Timer :} [8] tracks escalating conflict
    ℹtem \textbf{Stakes:} Control of psionic resources, civilian safety, political influence
\end{itemize}

\subsection{The Astral Breach}
\textbf{A rift between realms is growing larger}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Minor Threats:} 4−6 encounters with astral creatures
    ℹtem \textbf{Moderate Threats:} 2−3 encounters with reality distortions
    ℹtem \textbf{Major Threat:} 1 encounter with entity trying to cross over
    ℹtem \textbf{Campaign Timer :} [6] tracks breach expansion
    ℹtem \textbf{Stakes:} Reality stability, psionic safety, dimensional balance
\end{itemize}

\subsection{The Memory Plague}
\textbf{A psychic disease is spreading through telepathic contact}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Minor Threats:} 3−5 encounters with infected individuals
    ℹtem \textbf{Moderate Threats:} 2−3 encounters with disease vectors
    ℹtem \textbf{Major Threat:} 1 encounter with plague origin/source
    ℹtem \textbf{Campaign Timer :} [6] tracks disease spread
    ℹtem \textbf{Stakes:} Mental health, social stability, psionic community survival
\end{itemize}

\subsection{The Precognition Paradox}
\textbf{Future visions are becoming contradictory and dangerous}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Minor Threats:} 2−4 encounters with time−displaced effects
    ℹtem \textbf{Moderate Threats:} 2−3 encounters with fate manipulation
    ℹtem \textbf{Major Threat:} 1 encounter with entity controlling temporal flow
    ℹtem \textbf{Campaign Timer :} [8] tracks timeline instability
    ℹtem \textbf{Stakes:} Free will, destiny, causality itself
\end{itemize}

\section{Psionic Adventure Seeds}

\subsection{"The Silent Auction"}
\textbf{A black market auction for psionic artifacts and services}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Hook:} PCs need rare component/service only available at auction
    ℹtem \textbf{Location:} Secret venue with psychic security
    ℹtem \textbf{Complications:} Multiple bidders, artifact activation, betrayal
    ℹtem \textbf{Threats:} Corrupted psions, psychic predators, law enforcement
    ℹtem \textbf{Reward:} Unique artifact, valuable information, powerful contact
\end{itemize}

\subsection{"The Mind Prison Break"}
\textbf{Dangerous psychic criminals are escaping from specialized facility}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Hook:} PCs hired to recapture escapees or prevent breakout
    ℹtem \textbf{Location:} High−security psionic prison with mental wards
    ℹtem \textbf{Complications:} Guards compromised, reality distortion, public panic
    ℹtem \textbf{Threats:} Escaped prisoners, facility malfunction, competing interests
    ℹtem \textbf{Reward:} Substantial payment, criminal network disruption, public safety
\end{itemize}

\subsection{"The Astral Investigation"}
\textbf{Mysterious events require investigation in non−physical realms}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Hook:} Physical world problem with astral root cause
    ℹtem \textbf{Location:} Various planes of existence, psychic landscapes
    ℹtem \textbf{Complications:} Different rules, time distortion, identity confusion
    ℹtem \textbf{Threats:} Astral predators, trapped souls, reality anchors
    ℹtem \textbf{Reward:} Unique knowledge, interdimensional contacts, reality stability
\end{itemize}

\subsection{"The Emotion Plague"}
\textbf{Contagious emotional state is spreading through population}
\begin{itemize}[leftmargin=*]
    ℹtem \textbf{Hook:} PCs immune or need to stop spread for personal reasons
    ℹtem \textbf{Location:} Urban environment with high population density
    ℹtem \textbf{Complications:} Mass hysteria, social breakdown, economic impact
    ℹtem \textbf{Threats:} Empathic manipulators, infected crowds, authority response
    ℹtem \textbf{Reward:} Social influence, economic opportunities, emotional mastery
\end{itemize}

\section*{Conclusion}

This bestiary and generator system provides flexible tools for creating diverse psionic encounters that maintain the unique flavor and mechanical identity of the Fate's Edge psionics system while offering GMs quick, customizable options for campaign play. The integration of consequence tracks, SB generation, and psionic−specific mechanics ensures that every encounter reinforces the core themes of mental discipline and its inherent risks.

Use these generators to populate your campaigns with memorable psychic threats, create dynamic environments that challenge psionic characters in unique ways, and develop ongoing story threads that explore the deeper implications of mental power in your game world.

Remember that the true strength of psionics lies not just in its mechanical benefits, but in the narrative possibilities it creates. Every psychic encounter should ask questions about the nature of consciousness, the responsibility of power, and the thin line between mastery and madness.

\section{Psionic Investigation Procedures}

The unique capabilities of psionic characters have created new methodologies for investigation and detection that integrate seamlessly with Fate's Edge's existing investigative frameworks. These procedures transform how mysteries are solved while maintaining the core tension between information gathering and consequence management.

\subsection{Mental Detective Work}

Psionic investigators operate within the same framework as traditional investigators but with enhanced capabilities that come with their own risks and limitations. The key is that psionic abilities don't eliminate the need for traditional investigation skills−−−they amplify and complicate them.

⫕section{Memory Reading and Reconstruction}
Telepaths can access witness memories, but this process is far from simple information extraction:

\textbf{Memory Integrity Test:} When reading a witness's memory, the psion must make a Wits + Psionics vs. DV 3−5 (based on memory clarity, emotional trauma, and time elapsed). 
\begin{itemize}
    ℹtem \textbf{Success:} Gain Clue +1 and can ask one specific question about the memory's context
    ℹtem \textbf{Partial:} Gain Clue +1 but the memory is fragmented or confused with other events
    ℹtem \textbf{Miss:} Create False Memory [4] − the witness now "remembers" something that isn't true
\end{itemize}

\textbf{Mental Strain Cost:} Each memory reading costs 1 Mental Strain, plus additional Fatigue equal to the witness's Spirit attribute (representing their mental resistance).

⫕section{Deception Detection}
Psionic abilities enhance but don't guarantee the detection of lies:

\textbf{Surface Thoughts Scan:} Wits + Psionics vs. DV 3 to read immediate emotional reactions during questioning.
\begin{itemize}
    ℹtem \textbf{Success:} Identify general emotional state (truth, deception, fear, confusion)
    ℹtem \textbf{Partial:} Detect emotional inconsistency but cannot determine its source
    ℹtem \textbf{Miss:} GM may spend SB to create False Confidence − the psion becomes convinced of truth when deception exists
\end{itemize}

\textbf{Deep Intent Reading:} Spirit + Psionics vs. DV 4−5 to probe underlying motivations and truthfulness.
\begin{itemize}
    ℹtem \textbf{Cost:} 2 Mental Strain plus target's Presence (if willing) or Spirit (if resistant)
    ℹtem \textbf{Risk:} On miss, tick Psychic Backlash [4] − the target's mental resistance creates lasting strain
\end{itemize}

\subsection{Crime Scene Psionics}

Certain psionic abilities can extract information from crime scenes, but this practice is heavily regulated in most jurisdictions:

⫕section{Residual Impression Reading}
Very sensitive telepaths can sometimes read "echoes" left by intense emotional or traumatic events at crime scenes.

\textbf{Procedure:} Wits + Psionics vs. DV 4−6 (based on time elapsed, intensity of emotions, and environmental interference)
\begin{itemize}
    ℹtem \textbf{Success:} Gain vivid impression of final moments, potentially identifying perpetrator or victim state
    ℹtem \textbf{Partial:} Fragmented impressions that require interpretation and may be misleading
    ℹtem \textbf{Miss:} GM spends SB to create Psychic Contamination [4] − the psion absorbs traumatic elements that affect future investigations
\end{itemize}

\textbf{Legal Implications:} In most civilized areas, residual impression reading requires:
\begin{itemize}
    ℹtem Official authorization from investigative authorities
    ℹtem Presence of legal witnesses during the procedure
    ℹtem Documentation of all impressions in official records
    ℹtem Potential Psychic Evidence [6] timer that tracks admissibility in court
\end{itemize}

\subsection{Psionic Evidence and Legal Complications}

The introduction of mental evidence has created significant legal and ethical challenges across different jurisdictions:

⫕section{The Admissibility Problem}
Most legal systems struggle with how to treat information obtained through psionic means:

\textbf{Aeler Courts:} Require dual testimony − both the psion's interpretation and traditional evidence must support conclusions. Mental evidence alone is inadmissible.

\textbf{Lethai Moots:} Accept mental evidence when proper context witnesses can verify the psion's reading, but require extensive documentation of the context keys used.

\textbf{Ykrul Bowl Proceedings:} Mental evidence is acceptable if obtained without deception and with the subject's implied consent (no explicit refusal).

\textbf{Free Haven Practice:} Generally more permissive but require the psion to testify about their methods and potential for error.

⫕section{Chain of Custody for Mental Evidence}
Just as physical evidence requires careful handling, mental evidence must maintain "psychic chain of custody":

\begin{enumerate}
    ℹtem \textbf{Initial Reading:} Documented with time, location, and witness presence
    ℹtem \textbf{Storage:} Mental impressions must be either immediately recorded or maintained in the psion's memory with regular verification
    ℹtem \textbf{Transfer:} If transferring evidence to another psion, both must witness the transfer and document any changes
    ℹtem \textbf{Presentation:} In court, the psion must recreate the evidence from memory, with potential for Mental Strain costs
\end{enumerate}

\subsection{Investigation Integration with Other Modules}

Psionic investigation procedures naturally connect with other Fate's Edge systems:

⫕section{Political Intrigue Integration}
In political investigations, psionic abilities create both opportunities and risks:
\begin{itemize}
    ℹtem Reading political opponents' true intentions
    ℹtem Detecting treason or hidden loyalties
    ℹtem Risk of political retaliation if discovered
    ℹtem Potential for creating false evidence through Psychic Manipulation [4] timers
\end{itemize}

⫕section{Horror Campaigns Enhancement}
Psionic investigation in horror settings amplifies both discovery and danger:
\begin{itemize}
    ℹtem Reading the memories of victims or witnesses increases Dread Timer segments
    ℹtem Encountering alien or corrupted minds may cause Mental Corruption [4]
    ℹtem Precognition abilities may reveal horrific truths that increase Reality Fracture Timer 
\end{itemize}

⫕section{Caravan and Trade Investigation}
Psionic abilities are invaluable for detecting smuggling, fraud, and commercial espionage:
\begin{itemize}
    ℹtem Reading merchants' true intentions during negotiations
    ℹtem Detecting hidden cargo or false manifests
    ℹtem Risk of Merchant Guild retaliation for unauthorized mental intrusion
    ℹtem Potential for creating Business Intelligence [6] assets through mental reconnaissance
\end{itemize}

\subsection{Professional Psionic Investigators}

Specialized investigators have emerged who focus primarily on psionic methods:

⫕section{The Memory Advocates (Guild)}
Operating primarily in legal systems that accept mental evidence, these investigators combine traditional detective work with psionic enhancement:
\begin{itemize}
    ℹtem \textbf{Services:} Witness memory recovery, deception detection, residual impression reading
    ℹtem \textbf{Limitations:} Require legal authorization, maintain extensive documentation, face professional liability for false evidence
    ℹtem \textbf{Costs:} 2−5 times normal investigation fees plus potential legal complications insurance
\end{itemize}

⫕section{The Silent Seekers (Freelance)}
Independent investigators who operate in jurisdictions with less regulation:
\begin{itemize}
    ℹtem \textbf{Services:} Covert mental investigation, missing person location through memory tracing, corporate espionage detection
    ℹtem \textbf{Risks:} Legal vulnerability, potential for Mental Burnout [6] from overuse, reputation−based economy
    ℹtem \textbf{Network:} Maintain informal connections with law enforcement who use their services unofficially
\end{itemize}

These investigation procedures ensure that psionic characters remain meaningfully engaged in investigative scenarios while preserving the core tension and complexity that makes Fate's Edge investigations challenging and interesting. The integration maintains mechanical balance by ensuring that psionic advantages come with proportional costs, risks, and complications that create ongoing story opportunities rather than simple problem−solving.

\section{Racial Psionic Integration}

The development of psionic abilities has created unique relationships between different peoples and mental powers, with each culture's existing strengths and limitations shaping how they approach and utilize these extraordinary capabilities.

\subsection{Aeler: The Deep Accountants}

The Aeler's existing focus on precise measurement and resource management has profoundly influenced how they develop and utilize psionic abilities. For the Aeler, psionics represent another form of infrastructure that must be carefully maintained and accounted for.

⫕section{Breath Accounting and Mental Strain}
Aeler psions experience their mental strain through the same breath−debt system that governs all underground activity:
\begin{itemize}
    ℹtem Each psionic use costs "breaths" that must be repaid through rest and proper ventilation
    ℹtem Underground environments increase Mental Strain costs by 1 segment due to thin air
    ℹtem Aeler holds maintain "breath banks" − communal rest chambers where psions can recover mental energy
\end{itemize}

\textbf{Deep Accounting Integration:} Aeler psions must maintain detailed ledgers of their mental activities, recording:
\begin{itemize}
    ℹtem Mental Strain expenditure and recovery times
    ℹtem Effectiveness of different abilities in various environments
    ℹtem Cost−benefit analysis of psionic vs. traditional approaches
\end{itemize}

⫕section{Infrastructure Psionics}
Aeler psions often specialize in abilities that enhance their society's infrastructure:
\begin{itemize}
    ℹtem \textbf{Structural Sense:} Enhanced ability to perceive stress points and weak spots in buildings and constructions
    ℹtem \textbf{Vent Reading:} Advanced understanding of air flow patterns that enhances telepathic range underground
    ℹtem \textbf{Resonance Mapping:} Ability to sense the "health" of infrastructure systems through mental connection
\end{itemize}

\textbf{Cultural Integration:} Aeler psions are highly valued for maintaining underground settlements, but their "breath debt" makes them dependent on community support. This creates a symbiotic relationship where psions provide essential services while requiring significant community investment in their maintenance.

\subsection{Lethai: The Context Keepers}

The Lethai's context−saturated communication system has created a unique approach to psionics where mental abilities are as dependent on social and environmental context as their spoken language.

⫕section{Context−Saturated Telepathy}
Lethai psions must navigate the same context−saturated communication as their non−psionic kin:
\begin{itemize}
    ℹtem Reading minds without proper "context keys" (place, season, relationship) produces garbled or misleading information
    ℹtem Mental communication requires the same ritualized preparation as spoken communication
    ℹtem Different mental "dialects" exist between Lethai−al, Lethai−ar, and Lethai−thora communities
\end{itemize}

\textbf{Mark Integration:} Lethai psions often enhance their abilities through traditional Marks (tattoos, scars, resin−inlay) that serve as mental anchors:
\begin{itemize}
    ℹtem Spider's Patience Mark grants enhanced concentration for telepathic eavesdropping
    ℹtem Serpent's Breath Mark enhances ability to detect deception and hidden intentions
    ℹtem Storm−Shadow Mark allows for enhanced precognition during weather events
\end{itemize}

⫕section{Gift−Based Psionic Development}
The Lethai's Gift division creates distinct psionic development paths:
\begin{itemize}
    ℹtem \textbf{Body−Gift Psions:} Enhanced physical psionic abilities (telekinesis, biofeedback) that align with physical excellence
    ℹtem \textbf{Mind−Gift Psions:} Enhanced mental abilities (precognition, telepathy) that complement mental acuity
    ℹtem \textbf{Bridge−Born Psions:} Rare individuals who can combine both types but face social skepticism
\end{itemize}

\textbf{Cultural Integration:} Lethai psions serve as living archives and context−mediators, but their knowledge becomes "cursed" when removed from proper context. This creates a natural limitation that prevents them from becoming overpowered while maintaining their cultural significance.

\subsection{Ykrul: The Weather Readers}

The Ykrul's mastery of Kon'reh (sacred geometry) and their relationship with open spaces has shaped their approach to psionics around pattern recognition and tactical awareness rather than direct mental manipulation.

⫕section{Kon'reh−Augmented Precognition}
Ykrul psions don't see the future directly−−−they read the "flow/weight/exit" patterns of probability through their existing geometric understanding:
\begin{itemize}
    ℹtem Precognition is tied to route−finding and tactical awareness rather than personal fortune−telling
    ℹtem Mental visions appear as geometric patterns that must be interpreted through Kon'reh principles
    ℹtem Group precognition is more effective than individual − multiple psions can "map" probability spaces
\end{itemize}

\textbf{Weather Integration:} Ykrul psions have enhanced sensitivity to atmospheric conditions that affect their abilities:
\begin{itemize}
    ℹtem Storm conditions enhance telepathic range but increase Mental Strain costs
    ℹtem Clear weather allows for more precise precognitive readings
    ℹtem Wind patterns can carry or disrupt mental communications over long distances
\end{itemize}

⫕section{Tactical Psionics}
Ykrul psions excel in abilities that support their nomadic, clan−based society:
\begin{itemize}
    ℹtem \textbf{Herd Sense:} Enhanced ability to perceive the emotional and physical state of groups
    ℹtem \textbf{Route Reading:} Mental mapping that combines geographic knowledge with tactical awareness
    ℹtem \textbf{Banner Communication:} Long−distance mental communication between clan members
\end{itemize}

\textbf{Cultural Integration:} Ykrul psions are invaluable scouts and tacticians, but their "sightings" must be interpreted through Kon'reh principles. This creates a collaborative relationship where psions provide raw data that clan leaders must interpret through their cultural wisdom system.

\subsection{Vilakari: The Information Brokers}

The Vilakari's expertise in information networks and their position as cultural intermediaries has led to the development of psionic abilities focused on information gathering and social manipulation.

⫕section{Networked Telepathy}
Vilakari psions excel at surface thoughts and emotional currents in crowded spaces:
\begin{itemize}
    ℹtem Enhanced ability to read emotional undercurrents in marketplaces and social gatherings
    ℹtem Difficulty with deep, private thoughts due to cultural emphasis on surface interactions
    ℹtem Natural talent for detecting deception in commercial negotiations
\end{itemize}

\textbf{Thought−Markets:} Vilakari psions have developed "thought−markets" where information is traded through mental contact:
\begin{itemize}
    ℹtem Information brokers who specialize in mental exchanges
    ℹtem Standardized "thought−prices" for different types of information
    ℹtem Reputation systems based on the accuracy and usefulness of mental information
\end{itemize}

⫕section{Commercial Psionics}
Vilakari psions often specialize in abilities that enhance their information−brokering activities:
\begin{itemize}
    ℹtem \textbf{Surface Reading:} Enhanced ability to read immediate thoughts and intentions in social situations
    ℹtem \textbf{Memory Trading:} Ability to temporarily share specific memories for commercial transactions
    ℹtem \textbf{Negotiation Sense:} Enhanced perception of the underlying dynamics in business dealings
\end{itemize}

\textbf{Cultural Integration:} Vilakari psions operate underground information networks, but their abilities make them vulnerable to mental "audits" by authorities. This creates a cat−and−mouse dynamic where psions must constantly balance the value of their abilities against the risks of detection.

\subsection{Cross−Cultural Psionic Development}

The interaction between different racial approaches to psionics has created unique opportunities and conflicts:

⫕section{Cultural Exchange Programs}
Some communities have established formal programs for cross−cultural psionic training:
\begin{itemize}
    ℹtem Aeler academies that teach Lethai context integration
    ℹtem Lethai moots that explore Ykrul tactical applications
    ℹtem Ykrul clan gatherings that learn Vilakari information networks
    ℹtem Vilakari brokerages that adopt Aeler accounting precision
\end{itemize}

⫕section{Hybrid Abilities}
Individuals of mixed heritage or extensive cross−cultural training may develop unique hybrid abilities:
\begin{itemize}
    \textbf{Aeler−Lethai:} Psions who combine precise mental accounting with context−sensitive communication
    \textbf{Ykrul−Vilakari:} Psions who merge tactical precognition with information brokering skills
    \textbf{Lethai−Ykrul:} Psions who blend context−based telepathy with geometric pattern recognition
\end{itemize}

\textbf{Social Dynamics:} Different cultures have varying levels of acceptance for psionic abilities, creating complex social landscapes where psions must navigate not just their mental abilities but also the cultural expectations and limitations of their communities. This integration ensures that psionic characters remain meaningfully connected to the setting's cultural diversity while providing unique mechanical and narrative opportunities that reflect their peoples' distinct worldviews and capabilities.

\end{document}