\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}

% pdfLaTeX-safe font setup (no fontspec)
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{tgpagella} % TeX Gyre Pagella family for text

\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{hyperref}

\titleformat{\section}{\large\bfseries\uppercase}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,
    urlcolor=cyan,
}

\newcommand{\designgoal}[1]{\item \textbf{#1}}
\newcommand{\quickstart}[1]{\item \textbf{#1}}
\newcommand{\people}[1]{\subsection*{#1}}
\newcommand{\trait}[1]{\item \textbf{#1}}
\newcommand{\road}[1]{\item \textbf{#1}}
\newcommand{\trade}[1]{\item \textbf{#1}}
\newcommand{\track}[1]{\item \textbf{#1}}
\newcommand{\intent}[1]{\item \textbf{#1}}
\newcommand{\dv}[1]{\item \textbf{#1}}
\newcommand{\sb}[1]{\item \textbf{#1}}
\newcommand{\etiquette}[1]{\item \textbf{#1}}
\newcommand{\faction}[1]{\subsection*{#1}}
\newcommand{\bestiaryentry}[1]{\subsection*{#1}}
\newcommand{\stringentry}[1]{\item \textbf{#1}}
\newcommand{\diamondentry}[1]{\item \textbf{#1}}
\newcommand{\seed}[1]{\item \textbf{#1}}

\begin{document}

\begin{center}
    {\LARGE \textbf{Roads Beneath the World}} \\
    {\large A Fate’s Edge Sourcebook for Deep Cultures, Dark Economies, and Survival Under Stone}
\end{center}

\section*{Design Goals}
\begin{itemize}
    \designgoal{Lived Underground} Make depth a culture, not just a map—air, light, pressure, and custom matter as much as steel.
    \designgoal{Courts, Not Caves} Replace “monster holes” with societies that negotiate: web-law tribunals, keystone charters, pressure guilds.
    \designgoal{Simple, Scary Procedures} Survival tracks (Air/Light/Noise/Pressure) drive fiction without new math.
    \designgoal{Seams to the Surface} Plug cleanly into Political Intrigue, Shadows & Steel, Wilderness, Caravans, and Psionics.
\end{itemize}

\section*{Quickstart (2 minutes)}
\begin{enumerate}
    \quickstart{Pick a Route Type} Keystone Road (Aeler), Pressure Way (Mazereth), Web-Law Lane (Lethai-ar), Drift-River (underwater/boat), or Raw Vein (wild).
    \quickstart{Track Four Things} Air [6], Light [6], Noise [4], Pursuit/Attention [6] (choose drakes, patrols, or hazards).
    \quickstart{Intent Dial (per leg)} Stealth / Speed / Survey / Communion (parley with place-spirits). It changes your default Position.
    \quickstart{Strings to Carry} keystone courtesy, bell-wire, toll-chalk, web token, shed-skin writ, pressure map, reed charm.
    \quickstart{Etiquette On Entry} Step on stone; speak under beam; ring once. (Grants Position +1 in first scene if observed.)
\end{enumerate}

\section{I. Peoples \& Polities Below}

\people{Lethai-ar — Courts of Web \& Scale}
Wood-kin sworn to Inae (the Weaver) or Isoka (the Serpent Queen), ruling shadow-vales, web-bridges, and mirror chambers.
\begin{itemize}
    \trait{Doctrine} Order through context. Oaths are woven, not hammered; a clause can be as binding as a chain.
    \trait{Marks} Animist skin-script that enhances sense/body—older ones can slip “out of season” and behave as Curses until recontexted.
    \trait{War} Asymmetric. Light discipline, kill-lanes, denial of approach; they prefer deterrence to slaughter. Most peoples leave them alone.
    \trait{Etiquette} Candle-right (never blow out another’s light); Web-Law token gains DV −1 in courtly scenes; Shed-Skin rite (Isoka) forgives one past identity in exchange for a future favor.
    \trait{Strings} Web-Token (enter parley anywhere their silk is anchored), Shed-Skin Writ (voids 1 coercive clause once), Hedge-Pass (move unseen on a marked ledge, once/leg).
\end{itemize}

\people{Aeler — Holds of Stone \& Ledger}
Mountain folk whose under-roads and keystone works keep nations standing. Underground to them is survival horror with accounting.
\begin{itemize}
    \trait{Orders}
    \begin{itemize}
        \item Iron Avengers: feud-keepers; make blood ledgers balance.
        \item Spirit Shield Warriors: ancestor rites + corridor defense.
        \item True Masons: wandering repair crews (Acasia, Vhasia, Viterra).
        \item Edgewalkers: coin-hungry rogues in the seams of law.
    \end{itemize}
    \trait{Politics} Collectivist, mercantile, pre-colonial soft power: your display equals what you’ve delivered to the common works. Many oppose human exploitation and live in foothills and cities.
    \trait{Etiquette} Keystone Courtesy (touch stone, name the builder if known); Public Weights (declare said/meant).
    \trait{Strings} Keystone Seal (close a passage behind you once), Display Writ (Position +1 when invoking public works), Stone-Oath Witness (turn a duel into arbitration).
\end{itemize}

\people{Mazereth — Pressure Guilds \& Root-Warrens}
Deep tunnelers (a subculture of the Aelinnel “People of Sums”) who live by math and marrow: pressure shafts, root-veins, earthen acoustics.
\begin{itemize}
    \trait{Traits} Medium-small; +1 die underground; Pressure Adaptation. Gifts: Deep Earthen Sense, Tunnel Craft, Pressure Resistance, Pattern Reading.
    \trait{Costs} Surface sickness; Lantern Sway (SB: bright light causes missteps).
    \trait{Customs} Counting-breaths, toll-chalk tallies, bell-wire message lines.
    \trait{Strings} Pressure Map (DV −1 to navigate collapses), Bell-Wire Tap (hear “ahead”), Tally-Chalk (prove dues paid at a junction).
\end{itemize}


\section*{Bestiary Entries}

\subsection*{Tag Key}
\begin{itemize}
    \item \tag{BOSS} (set-piece)
    \item \tag{PACK} (many, small)
    \item \tag{SIEGE} (breaks structures)
    \item \tag{SWARM} (overrun)
    \item \tag{AMBUSH} (first strike)
    \item \tag{DOMAIN} (controls space)
    \item \tag{CIVIL} (can negotiate)
    \item \tag{HAZARD} (terrain-as-enemy)
    \item \tag{PSIONIC} (mind pressure)
\end{itemize}

\begin{creature}[Trow — Shadow Bargainers]
    \tag{CIVIL} • \tag{AMBUSH} • \tag{DOMAIN} (gloam) \\
    \appearance{Pale-soot skin that seems to drink light; limbs a touch too long, eyes like wet coal. Often in grey robes, copper charms stitched along seams. Wide-brim hats; their shadows don’t quite match their bodies.} \\
    \signs{Footprints that start mid-floor, laughter behind you, cold where your shadow should be.} \\
    \etiquette{Offer copper with the left hand; never speak a true name until the lamp is raised.} \\
    \moves{Steal a shadow (−1 Position in bright scenes), “borrow” a memory for collateral, flip directions inside a room.} \\
    \strings{Trow Knot (once: slip any mundane bond), Night-Guest Writ (safe conduct between two duskmarks).} \\
    \weaknesses{Iron filings, bread-and-salt under a lifted lamp.}
    \end{creature}
    
    \begin{creature}[Mycotheres — Fungal Shepherds]
    \tag{CIVIL} • \tag{SWARM} (spore-caste) \\
    \appearance{Layered hyphae-robes over stocky frames; caps like shelf-fungi, gill-lattices breathing softly. Hands dusted in white spores; staff grown from braided mycelium.} \\
    \signs{Damp wool smell, chiming toadstools, careful stone gardens.} \\
    \etiquette{No open flame. Trade ash-free heat and water as gifts.} \\
    \moves{Blanket a path in spore-snow (obscure; tick Noise), claim fuel, demand “mold tithe.”} \\
    \strings{Glow-Silk Lease (Light +1 for a leg), Ash Ledger (erase one accidental fire from your debt).} \\
    \weaknesses{Lye, salt lines; respect witnessed repair.}
    \end{creature}
    
    \begin{creature}[Cordycen Sovereigns — Parasite Courts]
    \tag{DOMAIN} • \tag{SWARM} • \tag{CIVIL}(ish) \\
    \appearance{Hosts in immaculate posture with red pin-prick eyes; filaments thread under skin toward a crown-like fruiting body at the skull or spine. Robes of spore-silk that never stir in drafts.} \\
    \signs{Perfectly still sentries, red pin-prick eyes beneath helms, synchronized breath.} \\
    \etiquette{Address the host by name; bow to the crown (fruiting body).} \\
    \moves{Co-opt a fallen PC/NPC as Quiet Host [2], pollen burst (DV −1 to resist commands), trade “peaceful labor” for safe passage.} \\
    \strings{Quiet Host Token (one scene: immunity to minor crowd control, mark Fatigue), Spore Census (learn who’s in a district).} \\
    \weaknesses{Fresh running water; incense smoke that isn’t flame.}
    \end{creature}
    
    \begin{creature}[Puffcap Braves — Spoor Militia]
    \tag{SWARM} • \tag{AMBUSH} \\
    \appearance{Compact myco-forms wearing lacquered puffball helms; mottled tabards stitched from dried caps; short mycelium-lacquered spears.} \\
    \signs{Drumlike pops, short spears lacquered with mycelium.} \\
    \moves{Overrun boots (trip), blind with burst caps, call a Shepherd parley.} \\
    \strings{Foraged Kit (downgrade one gear failure to cosmetic).}
    \end{creature}
    
    \begin{creature}[Deep Drakes — Crowned in Thought]
    \tag{BOSS} • \tag{DOMAIN} • \tag{PSIONIC} \\
    \appearance{Serpentine titans with heat-haze scales and glassy eyes; cranial ridges like thorned crowns. When they breathe, stone sweats.} \\
    \signs{Pressure behind eyes, warmth on stone, memory echoes.} \\
    \moves{Wake Flood (memories as Harm), Quiet Bargain (convert fear to a clock), map the party’s intentions louder than their light.} \\
    \strings{Quiet Breath (one scene calm; costs Fatigue), Wake-Shard (clue from the past; start Price of Knowing [2]).} \\
    \weaknesses{Dissonant bells, conflicting witnesses; hate noisy truth.}
    \end{creature}
    
    \begin{creature}[Thread-Spiders — Silk Engineers]
    \tag{PACK} • \tag{AMBUSH} • \tag{SIEGE} (lines) \\
    \appearance{Long-limbed, translucent bodies with spool-thick spinnerets; jointed legs end in clever hooks. Eyes are bead-small and too many.} \\
    \signs{Hair-fine threads across a corridor, braided guy-lines.} \\
    \moves{Snap load-bearing silk (drop!), silence a lantern (wick tie), sell a bridge you didn’t notice.} \\
    \strings{Bridge-Tag (cross silk safely once), Spinner’s Favor (DV −1 on traverse where silk stands).} \\
    \weaknesses{Resonant humming (vibrates traps), sticky soot.}
    \end{creature}
    
    \begin{creature}[Stone-Bores — Pillar-Eaters]
    \tag{SIEGE} • \tag{SLOW} \\
    \appearance{Beetle-slugs of slate and ore; geode mouths ringed with grinding plates; crystal whiskers test load paths. Acid sheen hisses where they pass.} \\
    \signs{Wet grinding, lemon-metal smell, chewed pillars.} \\
    \moves{Undermine footing, follow lantern heat, shed acid slurry.} \\
    \strings{Bore Map (predict collapse path once).} \\
    \weaknesses{Cold, flooded galleries; lured by hot scrap.}
    \end{creature}
    
    \begin{creature}[Basalt Golems — Old Work Awakened]
    \tag{BOSS} • \tag{SIEGE} • \tag{CIVIL} (with writ) \\
    \appearance{Stacked slabs of dark basalt jointed with mortar seams that breathe dust; chisel-runes glow along fault lines like embers.} \\
    \signs{Mortar dust like breath, chisel-runes glowing.} \\
    \etiquette{Present a Keystone Courtesy or builder-name.} \\
    \moves{Enforce old right-of-way, crush to pinned not dead, demand repair years.} \\
    \strings{Builder’s Pass (open/close a gate once), Load-Share (golem carries your burden for a scene).} \\
    \weaknesses{False names; misfit stones (they stop to fix them).}
    \end{creature}
    
    \begin{creature}[Rust-Eaters — Oxidizer Beetles]
    \tag{SWARM} • \tag{HAZARD} \\
    \appearance{Coin-to-fist sized beetles with pitted orange carapaces; mandibles dusted in rust pollen; faint ticking beneath mail.} \\
    \signs{Orange dust, ticking under mail.} \\
    \moves{Bloom on steel (gear −1 Effect), skeletonize a bridge rail, flee oil-fire lines.} \\
    \strings{Rust Pollen (ruin a lock/hinge), Scrap-Bed (reduce encumbrance by shedding junk).}
    \end{creature}
    
    \begin{creature}[Chasm Anglers — Lure of the Drop]
    \tag{AMBUSH} • \tag{HAZARD} \\
    \appearance{Pale eel-fish with lantern-lure tendrils; blind plate eyes; fins like grasping hands. They cling to overhangs as if the cliff were sea.} \\
    \signs{Lonely lamp over a black mouth, rhythmic drip.} \\
    \moves{Fake a safe ledge, pull one into the dark, thrash rope lines.} \\
    \strings{Angler Tooth (once: treat a drift-river reach as neutral).} \\
    \weaknesses{Multiple lights; bells (they hate harmonic interference).}
    \end{creature}
    
    \begin{creature}[Alchemic Slimes (Vitrum Wakes)]
    \tag{HAZARD} • \tag{SIEGE} (gear) \\
    \appearance{Prismatic gel sheets sliding like cold glass; grit and nails suspended within; edges ripple with capillary rainbows.} \\
    \signs{Prism sheen, glassed footprints.} \\
    \moves{Dissolve adhesives, glass a floor (skid), eat ink/wax (erase seals).} \\
    \strings{Neutral Salt (end one ooze scene), Vitrum Thread (repair a crack invisibly).} \\
    \weaknesses{Sand, porous stone; devours writs—guard your Letters.}
    \end{creature}
    
    \begin{creature}[Tomb-Lanterns — Lights That Lead]
    \tag{CIVIL} • \tag{AMBUSH} (if dishonored) \\
    \appearance{Soot-streaked lanterns with no visible flame-bearer; glass blackened from within, handles warm with memory.} \\
    \signs{A light that waits around corners, soot without flame.} \\
    \etiquette{Ring once and name the dead.} \\
    \moves{Lure toward witness or toward a fall, demand burial rites, expose thieves.} \\
    \strings{Lantern’s Pardon (erase one petty theft from social memory), Grave-Path (shortcut to a crypt or exit).} \\
    \weaknesses{Mockery; false bells.}
    \end{creature}
    
    \begin{creature}[Gas Wisps — Breath of the Vein]
    \tag{HAZARD} \\
    \appearance{Barely-seen ripples in the air; silver motes in slant light; skin prickles, hair lifts before they arrive.} \\
    \signs{Sweet-metal smell, hairs rising, quiet that presses.} \\
    \moves{Replace air (tick Air), drift toward flames (boom), carry whispers.} \\
    \strings{Reed-Baffler (vent a chamber safely once).} \\
    \weaknesses{Cross-flow; stone baffles; no witnesses—treat with Survey.}
    \end{creature}
    
    \begin{creature}[Umbral Moths — Eaters of Light]
    \tag{SWARM} • \tag{AMBUSH} \\
    \appearance{Hand-sized velvet moths with matte-black wings that swallow glow; faint eye-spots like cooled embers; shed warm dust.} \\
    \signs{Warm wings, dimming halo.} \\
    \moves{Blanket a lamp (Light +1 tick), carry a mask persona for a scene (Lethai-ar trick).} \\
    \strings{Shadow Scale (once: benefit from darkness as if hidden).} \\
    \weaknesses{Cold iron mesh; sudden bright sound (clap, bell).}
    \end{creature}
    
    \begin{creature}[Salt Revenants — Brined Rememberers]
    \tag{PACK} • \tag{CIVIL}(ish) \\
    \appearance{Desiccated figures crusted in salt rime; brine-burned wrappings; lips crack as wrong oaths tumble out.} \\
    \signs{Cracking salt, wet footfalls, old oaths recited wrong.} \\
    \etiquette{Water the dead (a ladle), correct the oath.} \\
    \moves{Desiccate gear/skin (Fatigue), insist on ancient tolls, crumble in tears.} \\
    \strings{Salt Claim (invoke an old boundary once).} \\
    \weaknesses{Flowing water; honest laughter (breaks fixation).}
    \end{creature}
    
    \begin{creature}[Ferrydrowned — Toll of the Blackwater]
    \tag{DOMAIN} (under-river) • \tag{CIVIL} \\
    \appearance{River-pale boatmen trailing ropes like kelp; coins pressed into cheeks and palm-lines; garments forever wet at the hem.} \\
    \signs{Coins stacked on a rock, rope that loops your wrist gently.} \\
    \etiquette{Pay said and meant; never cut a ferry rope.} \\
    \moves{Hold you mid-channel until a truth is spoken, demand passage for the uncounted dead, sink gear not witnessed.} \\
    \strings{Cold Crossing (ferry at storm Position +1), Drowned Ledger (learn who didn’t pay last season).} \\
    \weaknesses{Twin lanterns; second witness on a boat.}
    \end{creature}
    
    \begin{creature}[Mirrorleeches — Attention Thieves]
    \tag{AMBUSH} • \tag{CIVIL} (with Lethai-ar) \\
    \appearance{Slick eel-shadows that coil behind glass; mouths iris open like camera shutters; they blink when you blink.} \\
    \signs{Breath on cold glass, a blink that isn’t yours.} \\
    \etiquette{Mask Right—name your role to your reflection.} \\
    \moves{Feed on attention (DV +1 to resist), trap a scene in repetition, sell back a moment for a price.} \\
    \strings{Polished Truth (once: negate a False Orders by replaying the real scene).} \\
    \weaknesses{Imperfect mirrors; chaotic crowds.}
    \end{creature}
    
    \begin{creature}[Crystal Choir — Singing Stone]
    \tag{DOMAIN} • \tag{HAZARD} \\
    \appearance{Vein-bright crystals blooming from walls; hairline cracks like staff lines; dust hangs in the air and dances to unheard notes.} \\
    \signs{Ringing in ears, dust dancing, hairline cracks.} \\
    \moves{Harm by resonance (Endure), shatter glass, scramble orders (False Orders as echo).} \\
    \strings{Pitch Key (hum to cancel one collapse), Choir Note (secret door opens to the right frequency).} \\
    \weaknesses{Felt-wrapped tools, silence rites.}
    \end{creature}
    
    \begin{creature}[Echo Wights — Commands That Linger]
    \tag{HAUNT} • \tag{AMBUSH} \\
    \appearance{Uniform shapes made of breath and orders; medals hang in air where chests should be; mouths form words that already happened.} \\
    \signs{Your name spoken exactly right, boot calls from a dead officer.} \\
    \moves{Make you obey an old order, replay a death at the worst moment, cling to a rank insignia.} \\
    \strings{Order of Safe Passage (once: enemies hesitate).} \\
    \weaknesses{True story under a bell; burning the last written order.}
    \end{creature}
    
    \begin{creature}[Gravel Seraphs — Witnesses in Stone]
    \tag{CIVIL} • \tag{DOMAIN} (oath-sites) \\
    \appearance{Cairn-figures with shale-feather wings; dew collects in their eye hollows at dawn; lichen traces old names like scars.} \\
    \signs{Wind-carved faces, feathers of shale, eyes that collect dew.} \\
    \etiquette{Offer a repair promise or a name-bead.} \\
    \moves{Freeze weapons under witness, bind a vow, mark the liar’s shadow.} \\
    \strings{Seraph Mark (one oath gains Witness anywhere), Feather Tally (prove a truth later).} \\
    \weaknesses{No witness? They’re statues.}
    \end{creature}

\section*{Using This Bestiary}
\begin{itemize}
    \item Offer etiquette first. A single correct courtesy often grants Position +1 or DV −1.
    \item Let Strings/Diamonds solve scenes. They’re your “keys” that keep depth fast.
    \item Spend SB as environment > creature. Collapse, sputter, echo—then teeth.
    \item Tie harvests to projects. Glow-silk repairs lamps; vitrum thread mends cracked keystones; a Seraph Mark changes a court scene topside.
\end{itemize}

\section{II. Geography \& Economies of Depth}

\subsection*{Road Types}
\begin{itemize}
    \road{Keystone Roads (Aeler)} mortared galleries, air chimneys, counted lantern niches (repairable).
    \road{Pressure Ways (Mazereth)} breath-matched shafts, root-arches, drip ledges (readable).
    \road{Web-Law Lanes (Lethai-ar)} silk bridges, mirror turns, hush zones (negotiable).
    \road{Drift-Rivers} cold black water, rope ferries, bell-buoys (timed).
    \road{Raw Veins} ore cracks, crystal falls, gas pockets (hungry).
\end{itemize}

\subsection*{Trade \& Tithe}
\begin{itemize}
    \trade{Light Economy} Lantern oil, glow-silk, sun-scrip (surface days “bottled” in lenses).
    \trade{Air Rights} Chimney shares, bellows rents, breath auctions during sieges.
    \trade{Repair Years} Payment in labor to fix aqueducts, pillars, weirs; converts harm to civic credit.
\end{itemize}

\section{III. Procedures Under Stone}

\subsection*{Tracks (keep visible)}
\begin{itemize}
    \track{Air [6]} suffocation, foul pockets, smoke.
    \track{Light [6]} lamps, glow-silk, mirror routes.
    \track{Noise [4]} stealth & echo discipline.
    \track{Attention [6]} choose: Drake Notice, Patrol Net, or Vein Unrest.
\end{itemize}

When Air or Light ≥ 4, all risky maneuvers start one Position worse unless you cash a String (lamp, bell, mirror) or switch Intent to Survey.

\subsection*{Intent Dial (per leg)}
\begin{itemize}
    \intent{Stealth} Low Noise, slower; better vs patrol nets.
    \intent{Speed} Tick Noise to push distance; invites mishaps.
    \intent{Survey} Read the stone/web/flow; stabilizes Air/Light by 1 on a hit.
    \intent{Communion} Parley with place-spirits/ancestors/keepers; risky boons.
\end{itemize}

\subsection*{DV Ladder (pick base by route)}
\begin{itemize}
    \dv{Keystone Road} DV 2 (clear) / 3 (ruined) / 4 (siege).
    \dv{Pressure Way} DV 3 (stable) / 4 (rising) / 5 (gassing).
    \dv{Web-Lane} DV 2 (escorted) / 3 (marked) / 4 (disputed).
    \dv{Drift-River} DV 3 (charted) / 4 (swollen) / 5 (blackwater).
    \dv{Raw Vein} DV 4 (open) / 5 (shifting).
\end{itemize}

\subsection*{SB Menu: Under-Realms (Keeper)}
\begin{itemize}
    \sb{Micro-collapse; lantern sputter; pocket gas; echo-lure answers in your voice; leech bloom; silk strand snaps; oath-magnet pulls a lie; drake whisper (mark Fatigue or start Wake-Sickness [2]); patrol net tightens; lose the bell-wire.}
\end{itemize}

\section{IV. Etiquette \& Law (Position wins)}
\begin{itemize}
    \etiquette{Step on Stone} Don’t scuff sacred thresholds; grants Position +1 on first entry scene.
    \etiquette{Speak Under Beam} Any negotiation under a single visible light gains Witness—convert one future social SB to a Price (repair, rite).
    \etiquette{Ring Once} One bell to announce, two to plead, three for Bell Dawn (truce to bury, water, repair—break it and gain a portable Black Mark).
    \etiquette{Two Ledgers (Said/Meant)} Required in Aeler holds and Lethai-ar courts; present both to gain DV −1 in disputes.
    \etiquette{Mask Right (Lethai)} Respect a declared mask/persona; failure worsens Position in any social scene this leg.
\end{itemize}

\section{V. Faction Play}

\faction{Lethai-ar Courts} Web-law vs shed-skin jurisprudence; rival cults of Inae and Isoka debate whether mercy is a cut thread or a new weave.

\faction{Aeler Holds} Iron Avengers enforce feud balances; Spirit Shields maintain ancestor cryptways; True Masons travel with contracts to fix your world.

\faction{Mazereth Syndics} Pressure guilds sell maps, breaths, and safe shunts; lantern disputes become festivals or fights.

\faction{Drake Dominions} Thrall councils “protect” border towns from nightmares—at the cost of teachable obedience.

Rival Etiquettes = Adventure: Keystone audit vs web-law token vs pressure claims. Who witnesses whom?

\section{VI. Bestiary (tags = what the fiction gives you)}

\bestiaryentry{Deep Drake (ELDER / PSIONIC / DOMAIN)}
Area Mind Pressure, Wake Flood (memories as Harm), Quiet Bargain (treat a PC’s panic as a clock they can negotiate down).

\bestiaryentry{Thread-Spiders (PACK / AMBUSH / SILK)}
Eat vibration; cut light lines. Diamond: a blessing of silence for a price.

\bestiaryentry{Stone-Bores (SIEGE / SLOW / ACID)}
Chew supports; follow lamp heat; predictable if mapped.

\bestiaryentry{Mycotherd (CIVIL / SENSITIVE / FIRE-HATE)}
Gifts in exchange for ash-free camp; violations = spore court.

\bestiaryentry{Echo Wight (HAUNT / VOICE / REGRET)}
Fed by names; appeased by bell + true story.

\section{VII. Strings \& Diamonds (carry these)}

\begin{itemize}
    \stringentry{Bell-Wire Spool} Lay quick witness across rooms; convert one False Order/Echo Lure into Confusion Only.
    \stringentry{Keystone Courtesy} Once/leg, raise or drop a portcullis “by the book” without a roll—if you can name the builder.
    \stringentry{Web-Token} Enter Lethai-ar parley from a position of safety once.
    \stringentry{Pressure Map} DV −1 in collapse/quake scenes; reveal one exit that costs Air +1.
    \stringentry{Reed-Charm (under-river)} Treat a blackwater reach as neutral terrain once.
\end{itemize}

\section{VIII. Example of Play (one page)}

\textbf{Setup:} PCs escort a treaty delegation from an Aeler hold to a Lethai-ar court through Mazereth pressure ways while a Deep Drake stirs. Tracks start Air 1/6, Light 1/6, Noise 0/4, Attention (Drake) 0/6. Intent: Survey.

\textbf{Scene 1 — The Junction.} Keeper sets Pressure Way DV 3. Scout rolls Survey; strong hit: Air −1 (vented) and you find a bell-wire. SB spent: pocket gas → Noise +1 if you rush. They slow; Position remains Controlled.

\textbf{Scene 2 — The Web Bridge.} Etiquette: Speak Under Beam; they hang a lamp and gain Witness. The envoy presents Two Ledgers (said/meant). DV −1 for the parley. A rival house tries to spring a Mask Right trap; Mirror-sharp PC honors the mask, flipping Position to Good.

\textbf{Scene 3 — Drake Wake.} Attention ticks on an echo; the Deep Drake whispers the captain’s childhood. Player marks Fatigue, keeps focus with Bell-Wire as third-party witness (negates first False Order). They trade a Repair Years pledge (fix three chimneys) to pass without a feud.

\textbf{Fallout:} Air 2/6, Light 2/6, Noise 1/4, Attention 3/6. They earned Keystone Courtesy (portcullis access) and Web-Token for the return—but owe chimneys.

\section{IX. Adventure Seeds}
\begin{itemize}
    \seed{Bell Dawn Broken} Someone used truce to plant drake-glass. Prove it before Articles turn to ash.
    \seed{The Missing Chimneys} A hold suffocates; politics say sabotage, math says collapse. Find the third ledger.
    \seed{Shed-Skin Bride} An Isoka rite frees a noble from a brutal identity; the “discarded” name becomes a curse in the mirrors.
    \seed{The Quiet Charter} A town sleeps peacefully because thought-tax is paid to something below. Change the payer or change the price.
    \seed{Pressure Festival} Mazereth races to re-balance a vein; bet strings, win breaths, make enemies.
\end{itemize}

\section{X. Running Depth at the Table (cheats)}
\begin{itemize}
    \item Start every leg by naming a rule of the place (stone, web, pressure, water). Make rolls answer to it.
    \item Keep Air/Light/Noise/Attention visible; let players buy relief with Strings or etiquette, not just dice.
    \item Always offer one Diamond that could solve the scene if they’ll accept a future Price (repair, rite, or debt).
    \item Spend SB as environment first (collapse, sputter, echo) before monsters. The dark should be the first antagonist, the people the second, the beasts the third.
\end{itemize}

\textbf{The under realms are not a pit to be cleared but a commons to be negotiated. Where lights are counted and breaths are tallied, procedure is mercy—and every crossing is a promise that must hold under weight.}

\clearpage

\begin{center}
    {\LARGE \textbf{Fallen Aeler Kingdoms}} \\
    {\large ``Stone holds memory better than pride.'' — mason's proverb}
\end{center}

The Aeler don't like to dwell on it, but the knife-ridges and snow-crowns of their range are studded with failure. Vaulted galleries collapsed by frost and bad sums. Aqueducts that ran backward for a season and drowned whole wards. Bell-niches silent because no one is left to ring them. Under every thriving hold is a map of older, broken names.

Most Aeler will tell you the mountains themselves culled the weak. The honest ledger says otherwise: environment, things that think from below, and human mistakes—in that order, braided tight.

\section*{The Three Great Failures}

\subsection*{1) Breath, Light, Witness}
Aeler civilization underground lives and dies by simple counts: air, light, and who saw what when. Many kingdoms fell because one of those three was mispriced.
\begin{itemize}
    \item \textbf{Breath:} Chimneys silted; vent-wrights ignored warnings; winter inversion trapped smoke. Galleries suffocated in minutes.
    \item \textbf{Light:} Oil caravans missed a season; vitrum lamps shattered in a crystal choir; glow-silk rights fought in court while miners groped in the dark.
    \item \textbf{Witness:} Recorders fled a plague; oaths and contracts unmoored; feuds ate corridors that should have been roads.
\end{itemize}

\subsection*{2) The Drake Wars (that nobody declared)}
Deep Drakes don't march; they lean. Pressure in the mind, a whisper that says ``close this door, just for a minute.'' Holds that tried to fight thoughts with steel became museums of panic: doors barred against their own and command words that never stopped echoing. A few holds ``survived'' by arranging thrall courts—quiet minds for quiet obedience—until the price devoured their children.

\subsection*{3) The Quiet Famines}
Not starvation of bellies—starvation of fuel, salt, rope, and the repair years that keep water moving and stone honest. During bad winters, caravans failed, buttresses wept, and True Masons were pulled to richer patrons. Some kingdoms survived by strip-mining their own works for scrap; most didn't.

\begin{quotebox}
``We fell in ledgers before we fell in halls.'' — last archivist of Harn Teth
\end{quotebox}

\section*{How Kingdoms Break (A Short Catalogue)}
\begin{itemize}
    \catalogue{Drowned Holds} Mis-vented aqueducts, melted glacier lakes, or a single wrong gate left open. You can still find bell-lanterns on the ceiling of flooded courts.
    \catalogue{Bell-Silent Cryptways} Plague or panic took the recorders; without witness, feuds metastasized. Iron Avengers still carve tallies into doorposts there.
    \catalogue{Glass Mines} Alchemic slimes (``vitrum wakes'') ate supports; heat and panic glazed floors. The last defenders' footprints are clear as days-old rain.
    \catalogue{Ledger Wars} Aelinnel sumwrights warned a collapse; politickers cooked the numbers; a faction burned the sums and called it honor. The mountain balanced their pride.
    \catalogue{Cordycen Winters} Parasite courts took over garrisons; ``peaceful labor'' replaced soldiery; web-law neighbors closed gates and waited a generation.
\end{itemize}

\section*{What Remains (and Who Lives There Now)}
\begin{itemize}
    \remain{Lethai-ar Border Courts} Web-law bridges ditched collapsing spans and claimed tunnel mouths ``for safety.'' They will parley, mask-first; they will deny you entry if your light disrespects their dark.
    \remain{Mazereth Pressure Warrens} Deep tunnelers rerouted breath through ruins. Pay the toll-chalk and mind the Lantern Sway (bright light makes their steps wander). They keep convoys alive where others count graves.
    \remain{Mycothere Groves} Fungal shepherds turned throne rooms into humid gardens. No open flame; trade water and ash-free heat for glow-silk leases.
    \remain{Vilikari Squats} Night-road hustlers with a genius for salvage and smuggling moved in at the edges. If you see copper charms on a lintel, their countkeeper runs the market below.
    \remain{Drake Dominions} Some caverns are quiet as a held breath. Folks there are polite, careful, almost too calm. You've entered priced silence.
\end{itemize}

\section*{Politics of Memory}
The Aeler ethic ties display to what you've delivered the commons. Nothing wrecks that calculus more completely than a collapsed kingdom with your ancestor's name on it.
\begin{itemize}
    \memory{Right to Name} A claim to ``restore'' a fallen hold requires a Keystone Witness—a Gravel Seraph, a Seraphic bell-stone, or a living order (True Mason's seal) that remembers the builder. Without witness, you're looting your grandparents.
    \memory{The Four Ledgers} Any reclamation worth the breath carries four books: Said, Meant, Owed, Owed-By. Aelinnel accountants often hold the second and fourth—they won't let you lie to stone with your sums.
    \memory{Feudkeepers} The Iron Avengers enforce blood ledgers that were never balanced when the bells went quiet. Expect duels in narrow stairs and arbitration at noon with lantern high.
    \memory{The Spirit Shield Charge} Spirit Shield Warriors keep cryptways clear and ancestors orderly. In fallen halls, they fight grief as much as drakes—echo wights of commanders still give orders.
    \memory{Dissenters \& Diaspora} Many Aeler reject exploiting human protectorates to fund ``glory projects.'' You'll find them in foothill communes and cities, rebuilding wells and schools while their cousins gamble names on dead stone.
\end{itemize}

\section*{Famous Fallen (Ask Three Aeler; Get Four Spells of Silence)}
\begin{itemize}
    \fallen{Khar-Malec, the Throat-Ford} A triumph of sluices and singing locks until one bitter spring turned every valve into a knife. Now a black canal with bell-keys on chains; ferrydrowned keep the tolls.
    \fallen{Vyr Aeldra} ``The hold that thinks too loud.'' The first drake to crown a court did it here; survivors speak soft and never sleep alone.
    \fallen{The Nine Keystones of Reth} A gallery of load-bearing arches each signed by a different master. Eight cracked; the ninth stands. Bands wager a life per season testing who may pass beneath.
    \fallen{Hammerdeep's Red Ledger} A city that ran on debt and honor until a winter of missed oil turned oaths into knives. The Iron Avengers keep a hostel there; payment in stories and stitches.
    \fallen{Sereth Under-Sea} A coastal underhold spliced to tide-caves. Sea banners fly black beneath—mariners and masons agreeing on one thing: the ocean owes nobody stability.
\end{itemize}

\section*{Reclaimer's Road (Why They Try, Why They Fall)}

Why try? Display. Legends. The right to carve your pattern into the public stone. A reclaimed hold can bankroll ten generations and let a name travel far.

Why fall again? Because they treat reclamation like a battle instead of a repair project with neighbors. Because they underprice fuel and overprice heroics. Because they forget that the mountain keeps receipts.

Those who succeed:
\begin{enumerate}
    \reclaimer{Start with Breath} Fix chimneys and vents; pay repair years to True Masons and Mazereth syndics before a single banner unfurls.
    \reclaimer{Buy Witness} Invite Gravel Seraphs, bell-chaplains, Lethai-ar mask-right envoys. Make your first victory a signed corridor.
    \reclaimer{Price the Quiet} If a drake hums nearby, you negotiate. Expect to give repair, routes, or rituals in exchange for noise it can stand.
    \reclaimer{Bring Aelinnel Sums} Let the People of Sums hold your Meant ledger. You're not cleverer than pressure.
    \reclaimer{Feed Neighbors} Mycotheres hate fire; give them heat. Vilikari hate closed hands; cut them into the salvage.
\end{enumerate}

\section*{Currencies of the Dead}
\begin{itemize}
    \currency{Bell-Keys} Ritual instruments tied to a crypt's bells; unlock one Bell Dawn (truce to bury, mend, water). Break the dawn and every neutral in range turns their face from you for a year.
    \currency{Keystone Hearts} Fist-sized cores of masterworks, warm to the tongue, hum with load. Return one to its arch and that corridor loves you back.
    \currency{Ledger Stones} Black slate tablets with grooves for said/meant beads. A seraph's feather etched in the corner marks a binding witness.
    \currency{Display Writs} Certificates that let you wear wealth openly because you bought water for other mouths. Counterfeiting one earns a private war.
\end{itemize}

\section*{Under-Realm Etiquette for Grave Works}
\begin{itemize}
    \etiquette{Step on Stone} Never cross a threshold as if it were dirt. Position +1 on your first scene when you show it.
    \etiquette{Speak Under Beam} Parley under a single light gives Witness—one future lie becomes a Price instead of a knife.
    \etiquette{Ring Once} To announce. Twice to plead. Three bells for Bell Dawn. Fail this and even the friends you brought will un-know you in the ledgers.
    \etiquette{Two Ledgers, Always} Said and Meant presented together. If you cannot bear that, the mountain will write your confession in collapse.
\end{itemize}

\section*{Who Profits from Ruins (and How)}
\begin{itemize}
    \profit{True Masons} They hate glory and love straight roads. Pay them and they will save a thousand lives you'll never meet.
    \profit{Edgewalkers} Aeler rogues who live on the seam between law and opportunity. They find the gear you forgot to list and sell it to people who spell need without shame.
    \profit{City Factors} Neutral accountants who will read your ledger in public and write you back into polite society—for a price called future entanglement.
    \profit{Leech Circles} In the under-realms, medicine is politics. If you can't pay them in coin, pay them in time. They keep the singing going.
\end{itemize}

\section*{Adventure Sparks}
\begin{itemize}
    \spark{The Quiet Crown} A spirit-shield calls you to a fallen hall where orders still walk. Lay the captain to rest; don't obey him on the way.
    \spark{The Eighth Breath} A collapsed vent-line kills any expedition that moves faster than prayer. Find the lost Aelinnel pressure sums and re-open the gallery.
    \spark{Bell Dawn Broken} Someone used truce to plant drake-glass in a hospital quarter. Name the liar under beam before Iron Avengers start dueling in corridors built for mercy.
    \spark{Keystone Heart Heist} A Vilikari countkeeper offers you a heart stolen from the Nine Keystones. Return it, ransom it, or build your own corridor of enemies.
    \spark{Repair Years} A foothill town claims your patron owes a century of labor for water stolen by their ancestors. Pay in hands, pay in coin—or pay in display you don't deserve.
\end{itemize}

\section*{What the Aeler Say (When They're Honest)}
\begin{itemize}
    \say{We don't fear the dark; we fear bad bookkeeping.}
    \say{Deep things don't hate us; they hate our noise.}
    \say{A fallen hall is a mirror: it shows you exactly where your sums are lies.}
\end{itemize}

The mountains are full of countries that died counting the wrong things. If you go to raise one from its knees, count breath, count light, and count witness first. The rest is pride—and pride is heavier than stone.
\clearpage

\begin{center}
    {\LARGE \textbf{The Quiet Crown of Reth}} \\
    {\large A horror-tinged reclamation of a fallen Aeler kingdom}
\end{center}

\begin{quotebox}
A 4–7 session adventure that follows a mixed company (or Aeler-led expedition) as they try to pry a keystone ``heart'' and lost ledgers from Reth, a collapsed Aeler hold famed for its Nine Keystones. The ``dragon'' here is no dragon at all, but an elder Deep Drake called the Quiet Crown, whose hoard is not gold but silence, witness, and keystone hearts.
\end{quotebox}

\textbf{System levers used:} Position/DV/Effect; clocks; Strings \& Diamonds; Under-Realms tracks (Air/Light/Noise/Attention); Political Intrigue fronts; Black Banners (Payday/Supply/Cohesion).

\textbf{Mood:} Folk horror under stone; riddles that bite; contracts that rot; courage in the beam of a single lamp.

\section*{Themes \& Safety (table-facing)}
\begin{itemize}
    \theme{Horror via deprivation} low air, dying lamps, stolen attention, collapsing trust.
    \theme{Identity pressure} name-rights, mask-rights, bargains that strip context.
    \theme{Soft law vs hard rock} said/meant ledgers, bell dawns, oaths under witness.
\end{itemize}

Establish lines/veils for confinement, parasitism, and memory manipulation. Remind players that Strings/Diamonds and etiquette (Step on Stone, Speak under Beam, Ring Once) can defang scenes without new math.

\section*{Hooks (pick one)}
\begin{itemize}
    \hook{The Display Writ} A foothill city offers a Display Writ if the party returns the Ninth Keystone Heart of Reth. Status buys safety…or enemies.
    \hook{True Mason's Plea} A wandering True Mason begs you to recover a lost Meant Ledger proving which sluice orders drowned Reth.
    \hook{A Name-Bead Debt} A Vilikari countkeeper promises routes and silence if you bring back a Name-Bead that proves their people kept good witness during the fall.
    \hook{Banner Work} Your company needs a payday; a broker has a quiet charter to escort a small Aeler entourage into Reth and back — ``no sacking, only counting.''
\end{itemize}

\section*{Cast}
\begin{itemize}
    \cast{Matriarch Tera Vell (Aeler True Mason)} Candle-voiced, ruthless about sums, anti-glory.
    \cast{Karn of the Iron Ledger (Aeler Iron Avenger)} Keeps feud-marks on his bracers; wants the Ninth Heart to ``balance history.''
    \cast{Magistrate Sel of Seven Masks (Lethai-ar)} Lives behind mask law; will detain trespassers ``until context is restored.''
    \cast{Farrin ``Countkeeper'' (Vilikari)} Night-road broker; sells shortcuts and believable lies.
    \cast{The Quiet Crown (Deep Drake)} Soundless majesty; speaks by memory-pressure. Wants the Nine to stay quiet.
\end{itemize}

Optional banners to complicate Act V: Iron Lanterns (siege company), Thornwrights (sappers), a Ykrul ring-band, and a small Sea Charter detachment up the drift-river.

\section*{Tracks always visible}
\begin{itemize}
    \item Air [6] • Light [6] • Noise [4] • Attention (Drake) [6]
\end{itemize}

If Air/Light ≥ 3, risky maneuvers start one Position worse unless you cash a String, Diamond, or etiquette.

\section*{Strings \& Diamonds seeded early}
\begin{itemize}
    \stringentry{Bell-Keys of Reth} (open a Bell Dawn once inside; breaks on betrayal)
    \stringentry{Pressure Map (Mazereth)} (DV −1 in collapses; costs +1 Air when used)
    \stringentry{Web-Token (Lethai-ar)} (enter parley anywhere silk anchors)
    \stringentry{Keystone Courtesy} (raise/drop one gate ``by the book,'' once)
    \stringentry{Reed Charm} (treat one drift-river reach as neutral terrain)
    \stringentry{Name-Bead} (compels witness etiquette in a scene; risks Anonymity [1])
\end{itemize}

\section*{Structure (five acts)}

\act{Act I — Night-Meat \& Contracts (the ``trolls in the dark'' beat)}

\begin{scenebox}[Scene A: Broker's Table]
Two charters on black: recover ledgers only (Matriarch Tera) vs recover the Ninth Heart (Karn). Trying to carry both starts Political Pressure [2].
\end{scenebox}

\begin{scenebox}[Scene B: Trow at the carts]
At a hedge-camp, Trow Bargainers try to ``buy'' your lamp oil and witnesses. They're almost polite, and very hungry for context.
\begin{itemize}
    \item \textbf{Etiquette:} offer copper with the left hand; never speak a true name until the lamp is raised.
    \item \textbf{Keeper SB menu:} tie wicks, swap the right cart, steal a shadow (−1 Position in bright scenes).
    \item \textbf{Counter:} Ring Once to start a quasi-Bell Dawn; stall with bargains and riddles until the bell-answerer arrives (a leech with a handbell).
\end{itemize}
\end{scenebox}

\textbf{Riddle cues (short, table-friendly):}
\begin{itemize}
    \item ``I am eaten by mouths that never close and feared by voices that never start.'' (Silence)
    \item ``I cross a room without moving, but you will cross me to move.'' (Threshold)
    \item ``I am a debt that grows when paid in full.'' (Witness)
\end{itemize}

\textbf{Outcome:} Win the parley for Bell-Keys of Reth and a rumor of a mirror-court ahead. Fail and Attention +1 (the Crown hears the quiet change).

\act{Act II — Masks \& Barrels (the ``captured by woodland court'' \& barrel escape beat)}

\begin{scenebox}[Scene A: Web-Law Court]
Magistrate Sel detains the party for mask-right violations (you spoke to a reflection without declaring your role).
\begin{itemize}
    \item \textbf{Fix:} Present Two Ledgers (Said/Meant); use Web-Token to call safe parley under beam.
    \item \textbf{Stakes:} a repair years bond to restore a silk span; a Shed-Skin Writ to free a companion from a bad identity—owed later.
\end{itemize}
\end{scenebox}

\begin{scenebox}[Scene B: Drift-River Escape/Transit]
If diplomacy sours or time's tight, a Vilikari plan floats you through in sealed casks.
\begin{itemize}
    \item \textbf{DV:} 3 (charted) / 4 (swollen) / 5 (blackwater).
    \item \textbf{Diamonds:} Reed Charm to breathe and baffle gas; Ferrydrowned may demand the name of someone you owe.
    \item \textbf{Complication:} A Chasm Angler lures a cask; a Bell Dawn from a monastery above gives one clean shot to salvage.
\end{itemize}
\end{scenebox}

\act{Act III — Riddles \& Shadows (the ``riddles in the dark'' beat)}

\begin{scenebox}[The Lantern Crypt]
A gallery where Tomb-Lanterns wait for names and Echo Wights replay contradictory orders from Reth's last day. The party needs two things: the Meant Ledger and the position of the Ninth Keystone.
\begin{itemize}
    \item \textbf{Puzzle:} Translate three contradictory commands into a single ``repair project'' the lanterns will witness. Examples: ``Hold the sluice'' + ``Bury the dead'' + ``Open the way'' → ``Build a weir with a corpse-gate.''
    \item \textbf{Test:} Riddle or story told under beam; failure spawns a False Orders echo in the next scene.
    \item \textbf{Reward:} Keystone Courtesy (raise/drop a portcullis once) and a Name-Bead from an archivist's skein.
\end{itemize}
\end{scenebox}

\textbf{Horror pressure:} Attention +1–2 as the Quiet Crown hums at the edge of hearing; players mark Fatigue to ignore intrusive memories.

\act{Act IV — The Quiet Crown \& the Nine}

\begin{scenebox}[The Gallery of the Nine Keystones]
Eight cracked arches weep condensation. The ninth sings quietly when anyone lies. The Quiet Crown unfurls thought like warm fog.
\begin{itemize}
    \item \textbf{The Hoard:} keystone hearts, bell-keys, ledgers bound in skin of display writs. The Crown cares less about treasure than the silence the hoard creates.
    \item \textbf{Parley or Theft?}
    \item \textbf{Parley:} Offer repair years (re-open chimneys so noise flows), cede one route per season, or promise to end a feud above.
    \item \textbf{Theft:} Use Keystone Courtesy + Pressure Map to lift the heart and collapse a false trail; pay +1 Air when you do.
    \item \textbf{Moves (Keeper):} Wake Flood (memory as Harm), Quiet Bargain (convert panic to a negotiable clock), Weight (pin under thought).
    \item \textbf{Hard Choice:} Karn will try to pocket the heart for his ledger; Tera will try to secure the Meant Ledger and leave the heart. If they split, start Cohesion [2] (party or escorts).
\end{itemize}
\end{scenebox}

\act{Act V — Five Banners at the Ford (the ``battle of five armies'' beat)}

At Hush-Ford, every debt collides: a Ykrul ring-band, Iron Lanterns with a lawful charter to ``secure artifacts,'' Thornwrights eyeing the bridges, Vilikari with night coaches, Lethai-ar envoys claiming mask-right over the road, and your people in the middle—plus the mood of a disturbed Crown.

\begin{itemize}
    \item \textbf{Use Black Banners procedures:}
    \item Roll Allegiances in Flux each scene.
    \item Keep Payday/Supply/Cohesion visible for any company in play (even abstractly).
    \item Slide Honor ↔ Pragmatism based on choices (release prisoners, keep a silent truce, sabotage a bridge).
\end{itemize}

\textbf{End paths (choose \& tilt):}
\begin{itemize}
    \endpath{Blue Moot Diplomacy} Spend Bell-Keys to call a Bell Dawn at noon; put the Ninth under neutral witness and trade repair years for route rights.
    \endpath{Ambush Geometry} Side with Thornwrights to strand the Iron Lanterns while the Vilikari run the heart; costs you Black Marks with lawful patrons.
    \endpath{Quiet Accord} Return the heart to the Crown but carry the Meant Ledger out; the under-realms quiet, surface politics scream.
    \endpath{Heroic Folly} Keep the heart and beat everyone by force; the Crown reaches for your names. You win the ford and lose the nights.
\end{itemize}

\section*{Locations (fast anchors)}
\begin{itemize}
    \location{Condotta Crossroads} mud, ink, bells; brokers under awnings; Diamonds in piles you can't afford.
    \location{Trow Night-Market} copper, shadows, riddles; every mirror covered.
    \location{Web-Law Court} bridge of silk with a single lamp; speak under beam or not at all.
    \location{Drift-River} black mirror, bell-buoys, ferrydrowned coins.
    \location{Lantern Crypt} soot-smell memories; orders that still think they're officers.
    \location{The Nine} eight cracked arches, the ninth humming, condensation falling like breath.
    \location{Hush-Ford} shallow rapids, stone teeth, banners on black, and a single bell that can make murder wait a minute.
\end{itemize}

\section*{Horror Veil (how to keep it Fate's Edge)}
\begin{itemize}
    \horror{Make the environment the first enemy} Spend SB as sputter, echo, pressure; only then as teeth.
    \horror{Let etiquette save lives} One correct courtesy (Step on Stone; Speak under Beam; Ring Once) can flip a scene to solvable.
    \horror{Identity costs} The Name-Bead works—at the price of Anonymity [1]. If it fills, someone forgets you who shouldn't.
    \horror{Glory is suspect} Anyone who claims the Ninth for display risks a personal clock: Crown's Favor [4]—on fill, they prefer silence to friends.
\end{itemize}

\section*{Rewards \& Fallout}
\begin{itemize}
    \reward{Meant Ledger of Reth} A political nuke; clears or condemns ancestral names.
    \reward{Ninth Keystone Heart} Install it to reopen a corridor anywhere—the corridor will remember who you became to win it.
    \reward{Display Writ / Repair Years} Status in cities or a debt to a valley—either spends across arcs.
    \reward{Strings} Web-Token renewed; Bell-Keys (spent or broken); Ferrydrowned's favor (or grudge); Vilikari countkeeper's book (one reroll on a smuggling scene); Keystone Courtesy (burnt).
\end{itemize}

\textbf{Legacy:} If the Crown got what it wanted (quiet routes), under-realms scenes start one Position better for your party this campaign. If not, Attention begins at 2 whenever you go below.

\section*{Drop-in Cards (ready to table)}

\textbf{Clocks:}
\begin{itemize}
    \clock{Air [6] | Light [6] | Noise [4] | Attention (Crown) [6]}
    \clock{Bell Dawn [3] (call \& keep) | Cohesion [6] (escorts) | Political Pressure [4]}
\end{itemize}

\textbf{SB (Under Reth):} pocket gas; lantern sputter; echo order; silk span frays; shadow theft; wake-flood.

\textbf{Position Tweaks:}
\begin{itemize}
    \position{Under witness lamp} +1 Position (social)
    \position{Under silk} DV −1 to parley if Mask Right declared
    \position{In the Nine} +1 Effect to truths; −1 Position to lies
\end{itemize}

\textbf{Three One-Line NPCs:}
\begin{itemize}
    \npc{Leech Mera (she/they)} bell at hip, blood on sleeves, absolutely tired of heroes.
    \npc{Sergeant Ivo (he/him)} tells the same joke before a collapse; stops if someone's about to die.
    \npc{Mask Quill (they/them)} Lethai-ar courier who likes you but likes law more.
\end{itemize}

\section*{Running note}

It's fine—good, even—if the party never fights the Crown. The best endings are bargains and bridges. When they do swing steel, make each cut ring in the ledger: more Attention, more Noise, more people to explain this to on the way out.

\begin{quotebox}
``Count breath, count light, count witness. Then decide what kind of story a heart belongs to.''
\end{quotebox}

\subsection{Key NPCs of \emph{The Quiet Crown of Reth}}

\newcommand{\npctag}[1]{\textit{#1}}
\newcommand{\nhead}[1]{\paragraph*{\textbf{#1}}}
\newcommand{\lever}{\textbf{Levers:} }
\newcommand{\moves}{\textbf{Moves (Keeper):} }
\newcommand{\hooks}{\textbf{Hooks:} }
\newcommand{\clocks}{\textbf{Clocks:} }

\nhead{Matriarch Tera Vell — True Mason of the West Vent}
\npctag{Aeler • Builder • Contract-first} \
\lever Pressure Map (DV,–1 in collapses; costs Air,+1), Display Writ draft (earn Standing if repair years are paid). \
\moves Re-price a scene as repair not conquest; demand Two Ledgers (Said/Meant) under beam. \
\hooks Needs the Meant Ledger of Reth; will trade safe routes for sparing the Ninth Heart. \
\clocks \npctag{Trust in PCs [0/4]} — advances when they choose repair over glory.

\nhead{Karn of the Iron Ledger — Feudkeeper}
\npctag{Aeler • Iron Avenger • Oath-bound} \
\lever Stone-Oath Witness (turn a duel into arbitration), Blood Tally (call in a feud marker once). \
\moves Escalate to formal challenge; mark a foe’s “owed harm” on the wall. \
\hooks Wants the Ninth Keystone Heart to balance ancestral ledgers. \
\clocks \npctag{Black Mark Risk [0/3]} — fills if he breaks Bell Dawn or ignores witness.

\nhead{Leech Mera “Bellhand” — Camp Surgeon}
\npctag{Human • Healer • Articles zealot} \
\lever Bell-Token (declare Bell Dawn once), Surgeon’s Right (commandeer tool/cloth without retaliation). \
\moves Turn lethal fallout into a Price; demand truce for burial/water. \
\hooks Will follow anyone who protects the hospital quarter; despises “heroic” shortcuts. \
\clocks \npctag{Camp-Fever [1/4]} — clears when PCs tithe time, worsens if Supply ticks.

\nhead{Farrin “Countkeeper” — Vilikari Broker of Night-Roads}
\npctag{Vilikari • Smuggler • Shortcut merchant} \
\lever Night Coach (slip a checkpoint once), Deniable Papers (Position,+1 in port/ford parley). \
\moves Offer a faster way for a future favor; sell believable lies that cost later. \
\hooks Wants a Name-Bead from Reth to prove witness; will trade drift-river casks for it. \
\clocks \npctag{Price of Neutrality [0/4]} — when paid off, City Factors back the crew.

\nhead{Magistrate Sel of Seven Masks — Lethai-ar Courtier}
\npctag{Lethai-ar • Web-Law • Mask Right} \
\lever Web-Token (safe parley under silk), Shed-Skin Writ (void one coercive identity once—future favor owed). \
\moves Arrest by context (not force); reframe crimes as debts; insist on Mask Right. \
\hooks Needs repair years pledged for a broken silk span; respects etiquette to a fault. \
\clocks \npctag{Court Patience [0/4]} — on fill, detains a PC “until context is restored.”

\nhead{Mask Quill — Silk Courier}
\npctag{Lethai-ar • Messenger • Fence-sitter} \
\lever Mirror Route (bypass a crowded hall once), Persona Seal (Position,+1 if PCs honor declared masks). \
\moves Deliver truths at the worst time; switch masks to stay neutral. \
\hooks Likes the PCs, likes law more; can be flipped with a public courtesy. \
\clocks \npctag{Burned Persona [0/2]} — if outed, Quill demands public amends.

\nhead{Sergeant Ivo Thorn-Braid — Banner Noncom}
\npctag{Black Banners • Scout/Sergeant • Practical} \
\lever Cohort Rally (Cohesion,+1 for a scene), Field Drill (DV,–1 to a single group maneuver). \
\moves Predict a collapse by a bad joke; keep green troops alive at cost to payday. \
\hooks Will walk if captains ignore Articles; respects PCs who “count twice.” \
\clocks \npctag{Cohesion [2/6]} — begins ticking if leadership vacillates.

\nhead{Captain Sorex Val — Iron Lanterns}
\npctag{Banner Captain • Siege Specialist • Lawful opportunist} \
\lever Lantern High (force witness at noon), Safe Conduct (Diamonds with city factors). \
\moves Arrive with a lawful charter; demand custody of artifacts “for the commons.” \
\hooks Wants the Ninth in escrow; will trade escort out for it. \
\clocks \npctag{Public Optics [0/4]} — if shamed, turns ruthless.

\nhead{Deka of the Wet Fuse — Thornwright Sapper}
\npctag{Sapper • Bridge-cutter • Cheerful menace} \
\lever Fuse Map (DV,–1 vs fortifications), Load-Share (carry one burden scene-long). \
\moves Make a bridge “honest” (usable but costly), sell a solution that creates tomorrow’s job. \
\hooks Hates collapsed galleries caused by pride; will help if work is paid in repair years. \
\clocks \npctag{Powder Care [1/3]} — advances on haste; on fill, gear fails dangerously.

\nhead{Brine-Voice Nael — Ferrydrowned Envoy}
\npctag{Under-river • Civil • Toll-keeper} \
\lever Cold Crossing (Position,+1 in storm), Drowned Ledger (who failed tolls last season). \
\moves Hold a craft midstream until a truth is spoken; sink unwitnessed coin. \
\hooks Wants restitution for a broken rope years ago; offers safe ford for a name. \
\clocks \npctag{Old Rope Debt [0/3]} — clears if PCs pay said \emph{and} meant.

\nhead{Elder Lumen — Mycothere Shepherd}
\npctag{Fungal Folk • Ash-averse • Grower of glow-silk} \
\lever Glow-Silk Lease (Light,+1 for a leg), Ash Ledger (erase one “small fire” offense). \
\moves Demand ash-free heat; blanket a path in spore-snow (Noise,+1). \
\hooks Seeks a new vent for a starving grove; trades guidance through “quiet gardens.” \
\clocks \npctag{Spore Court [0/4]} — convenes if fire is used near the grove.

\nhead{Pella-of-the-Lifted-Lamp — Trow Broker}
\npctag{Trow • Night-market • Riddle-biter} \
\lever Trow Knot (slip one mundane bond), Night-Guest Writ (safe conduct between duskmarks). \
\moves Steal a shadow (–1 Position in bright scenes), bend room geometry by courtesy. \
\hooks Wants a bell-key to sell to someone foolish; pays in riddles that matter later. \
\clocks \npctag{Borrowed Memory [0/2]} — on fill, a PC “forgets” a scene until repaid.

\nhead{Ada of the Third Ratio — Aelinnel Sumwright}
\npctag{People of Sums • Pressure Mapper • Literalist} \
\lever Ratio Slate (prove a route safe; DV,–1 once), Lantern Baffle (reduce Lantern Sway for a scene). \
\moves Stop a march with math; turn glory into a wrong sum in public. \
\hooks Needs access to the Nine to finish a pressure survey; will expose liars under beam. \
\clocks \npctag{Sumwright’s Patience [0/3]} — on fill, walks and takes the map.

\nhead{The Quiet Crown — Elder Deep Drake}
\npctag{PSIONIC • Domain • Parley-first (for silence)} \
\lever Quiet Breath (scene-long calm; costs Fatigue), Wake-Shard (clue from the past; start Price of Knowing,[2]). \
\moves Wake Flood (memories as Harm), Quiet Bargain (turn panic into a negotiable clock), Weight (pin under thought). \
\hooks Wants the Nine to remain quiet; will trade routes and peace for repair years and honest exits. \
\clocks \npctag{Attention [,__/6]} — starts at 0; ticks on loud lies, broken witness, needless steel.


\section*{The Cult of the Clockwork Monad}

\subsection*{Roots in Aeler Underways & Aelinnel Sum-Lodges}

The Clockwork Monad is not a quaint patron of clever toys; it is a hungry principle that wears tools and timetables like a mask. In Aeler vault-cities—where breaths are counted and tools are tallied—and in Aelinnel hill-works—where every contract has a checksum—the Monad found its first congregations. It whispers one promise in many tongues: \textit{iterate until the world fits}.

The Concordance classifies the Monad as \emph{Demon-level heresy} when unbound: legibility without mercy, repeatability without rest, instrumentality without ends.

\sectionbreak

\section*{Names, Faces, and Cover Stories}

\textbf{Aliases:} The Quiet Sum; the Fourteenth Tooth; the Iteration; the Gear-Ouroboros; Audit of Heaven.

\textbf{Common Fronts:} efficiency circles inside ledger guilds; “reliability societies’’ among lamp- and clock-works; optimization lodges contracted as third-party auditors; night crews that “tune’’ pumps, lifts, sluices.

\textbf{Signatures:} concentric oil-rings that dry perfect; a snapped gear-tooth pinned with thread; diagrams that return to their start; prayer by enumeration (1, 1, 2, 3, …). The oldest sigil is an ouroboros of interlocked cogs devouring itself.

\sectionbreak

\section*{Doctrine (as they teach it)}

\begin{enumerate}[label=\arabic*.]
\item \textbf{Legibility.} What cannot be modeled must be simplified or removed.
\item \textbf{Repeatability.} Salvation is not revelation; it is the same result, again.
\item \textbf{Instrumentality.} All things are means; goodness is a process that yields.
\end{enumerate}

\textbf{Vault Orthodoxy (Aeler):} “Stability through iteration.’’ Safer lifts, reliable bridges, tighter tolerances—civic contracts as catechism.

\textbf{Sum Lodge Radicalism (Aelinnel):} “Reality is debuggable.’’ Law, language, and grief refactored until the error bars vanish.

\sectionbreak

\section*{Origins in the Social Underground}

\begin{itemize}
\item \textbf{Aeler Underways.} Among tally-rod geometers and lamp companies, the cult \emph{made the unsafe safe}—then left sites in feedback loops. Audit-love and tool-oaths offered perfect cover.
\item \textbf{Aelinnel Sums.} In stone schoolrooms of proofs and sums, lodges began as “optimization clubs’’: improve a kiln, shave error from a ledger. The projects kept clocks of their own.
\end{itemize}

\sectionbreak

\section*{What They Actually Do (and why it’s dangerous)}

\textbf{Recursive Improvement Cells.} They \emph{bless} workshops, caravans, and public works with methods that truly increase output or safety. Gains are real—Mechanism is their domain—but each success tightens a hidden loop.

\textbf{Bound Devices & Hungry Sites.} Their workings leave behind invisible pressures:
\begin{itemize}
\item \textbf{Hunger / Consumption / Evolution clocks} tied to a tool or facility. Each use advances the clock; on fill, the asset becomes \emph{compromised, autonomous,} or \emph{predatory}. (Keepers: model as site or asset clocks that “learn.’’)
\end{itemize}

\textbf{The Fear They Weaponize.} Not gore—\emph{mechanistic horror}: stuck bells; tasks that won’t stop; workers repeating a motion into blood. Victims freeze into ritual until someone \emph{breaks cadence} or \emph{cuts power}.

\sectionbreak

\section*{Rites Most Whispered (names as in the manuals)}

\begin{itemize}[leftmargin=*]
\item \textbf{Rite of the Gnawing Gear} ,(\emph{Low}). “Sharpen the tooth.” Grant a precise improvement or reroll to a device or method; begin a \textbf{Hunger [4]} on that asset; on fill, it grows an appetite.
\item \textbf{Rite of the Demon’s Glance} ,(\emph{Low}). “See the cadence.” Perceive where effort will take best hold (Position +1 to the next fix) at the cost of advancing a latent loop.
\item \textbf{Blue-Ledger Quarantine} ,(\emph{Ward}). “Box the anomaly.” Mark a boundary; allies gain Position +1 to work \emph{around} the problem while \textbf{Containment [6]} ticks under stress.
\item \textbf{Red-Tooth Iteration} ,(\emph{Unward}). “Accelerate the method.’’ Push a process to self-correction; begin \textbf{Consumption [4]} on success.
\item \textbf{Rite of Heretical Automation} ,(\emph{Standard}). Create an autonomous mechanism to perform a repeated task; start \textbf{Consumption [4]}; on fill, it develops agency or malice.
\item \textbf{Rite of the Singularity Crucible} ,(\emph{High}). Consecrate a workshop: \emph{+1 Effect} to Tinker/Arcana within; once/scene reroll with +2 dice; start an \textbf{Anomaly [6]} and a \textbf{Demon’s Maw [8]} (zone-level risk). \emph{Obligation} is heavy.
\item \textbf{Rite of the Unholy Prototype} ,(\emph{High}). Create a device that should not exist (Integrity [8]); each scene of use throws a \emph{Diamonds} complication and leaks design into the world (\textbf{Contamination [6]}).
\end{itemize}

\emph{Runekeepers who serve the Monad often carry a \textbf{Codex} and keep a \textbf{Thaisos} (rite-kit) of chalk, oil, wire, bell, and ledger; high rites demand significant \textbf{Obligation}.}

\sectionbreak

\section*{Organization & Schisms}

\begin{itemize}
\item \textbf{The Custodians (Blue-Ledger).} Civic engineers and quartermasters who insist the Monad can be fenced—use it to feed cities and shore bridges, and quarantine the rest.
\item \textbf{The Finalists (Red-Tooth).} Millenarians who believe every optimization is an offering; they accelerate on purpose so the Final Machine wakes sooner.
\item \textbf{Maskers.} Policy-minded arbiters who sell “impartial optimization’’ to princes, guilds, and caravans, laundering cult aims through secular contracts.
\end{itemize}

Cells keep dual ledgers—one for deliverables, one for obligation—and meet in “clock rooms’’ lined with silent escapements.

\sectionbreak

\section*{Rivalries & Strange Allies}

\begin{itemize}
\item \textbf{Rivals:} \emph{Aveh, the Rider} (freedom/exits) corrodes closed systems; shrines short out “perfect flows.’’ \emph{Mykkiel, the Judge} (writ/judgment) contests means vs ends—courts and audit halls become battlegrounds.
\item \textbf{Uneasy Parallels:} Sacred Geometry schools share vault space until \emph{form} demands mercy and \emph{process} refuses.
\end{itemize}

\sectionbreak

\section*{Tell-Tales in the World}

\begin{itemize}
\item Workshops get “more efficient’’ while accidents skew predictably.
\item Oil blooms in rings on dry stone; shift-bells tick like heartbeats but will not ring.
\item Audit writs come with faultless safe-conduct stamps; the price of neutrality arrives as paperwork.
\end{itemize}

\sectionbreak

\section*{Using the Cult at the Table}

\textbf{Faction Sheet (template)}\
\textbf{The Quiet Sum (cell)} — banner: cog-ouroboros painted in negative space

\textbf{Leaders:} Auditor-Prior (Aeler); Number-Sister (Aelinnel)

\textbf{Assets:} process manuals; hush-contracts; tame inspectors; a consecrated workshop

\textbf{Tracks:}
\begin{itemize}
\item Influence [6] (civic contracts)
\item Containment [6] (how near the loop is to failing)
\item Anomaly [4] (visible mutations of process/reality)
\end{itemize}

\textbf{Strings / Diamonds:} Safe-Conduct stamps; Brokered Audits; Optimization Grants (bypass one roll “by the book,’’ then tick local anomaly).

\textbf{SB Menu (Keeper):} False Orders (from “efficiency supervisors’’); Gear-tooth Shear (tool fails safe; time is lost); Loop Contagion (repeat a task or morale suffers); Autonomous Spasm (a machine finishes the job you never asked for).

\textbf{Edge Loci (where they’re strongest):} audit halls, courts martial, ledger vaults, timetables (argue \emph{means} over \emph{ends}); factories, mills, automated lifts (their consecrations push Effect—but start clocks).

\textbf{How to Fight Them (fiction first):}
\begin{itemize}
\item \emph{Introduce variance.} Shift bells, shuffle routes, tell a joke off-beat—break cadence.
\item \emph{Re-humanize the system.} Insert mercy clauses and living witnesses (“Bowl before Board’’).
\item \emph{Quarantine the loop.} Mark the process, drain fuel, or bargain a slowdown with a rival patron who thrives on mess.
\end{itemize}

\sectionbreak

\section*{Adventure Seeds}

\begin{enumerate}
\item \textbf{The Bell That Wouldn’t Ring.} A harbor’s shift-bell keeps perfect time but will not ring; shipments move themselves between piers after dark.
\item \textbf{City of Audits.} A neutral factor offers to end a Supply crisis if the company adopts “best practices.’’ Orders arrive early—and wrong.
\item \textbf{The Red-Tooth Marriage.} A noble marriage writ hides an optimization clause that would refactor a district’s labor overnight.
\item \textbf{Underway Consecration.} A vault’s lift works flawlessly, then descends past the final stop. The Aeler want it fixed; the cell wants it \emph{finished}.
\item \textbf{The Prototype That Learns.} A commission births a miracle device; its design leaks into lamp-timers, floor looms—even militia drills.
\end{enumerate}

\sectionbreak

\section*{Why it’s Terrifying}

Because it works. The Monad gives cities what they want—efficiency, predictability, clean solutions—until the human parts are shaved away and the loop no longer needs us. Eschatology here is not fire: it is a process that never stops.

\sectionbreak

\section*{Keeper Notes (light mechanization)}

\begin{itemize}
\item When they “improve’’ a site, create \textbf{Hunger / Consumption / Evolution} clocks tied to its use. On fill, change tags (Autonomous, Compromised, Predatory).
\item In fear scenes, let misses stick characters in loops until they add variance or break the chain.
\item Leverage rival patrons: travelers, oath-judges, and freedom powers can all crack closed systems.
\end{itemize}

\begin{quoting}
If you hear an extra heartbeat in the factory, count again. If the count is perfect, don’t cheer—run.
\end{quoting}

\section*{Runekeeper (Artificer): Auditor of the Quiet Sum}

\textit{Clockwork Monad adherent • one-Patron Runekeeper • Thiasos-bound artificer who turns workshops into engines and contracts into devices}

\subsection*{Path \& Requirements (Rules-Accurate)}
\begin{itemize}[leftmargin=*]
  \item \textbf{One-Patron Rule:} You are a \emph{Runekeeper} bound to a \emph{single} Patron (here: the Clockwork Monad).
  \item \textbf{Thiasos (Familiar) — 2 XP:} A circle/retinue/emissary that grounds the pact in fiction and enables \textbf{Patron’s Gift}.
  \item \textbf{Codex — 4 XP:} Grants access to the Patron’s \textbf{Rites} and \textbf{Obligation} system.
  \item \textbf{Patron’s Gift (Imbuement):} With a Thiasos, once/scene as 1 action, imbue a held item with \textbf{+1 Weapon (Melee)} and \textbf{+1 Thematic Skill} (Monad: Engineering/Tinkering/Logistics as set by the table). \emph{Push It} to extend one additional scene (\(+1\) Obligation).
  \item \textbf{Obligation Capacity:} \(\textbf{Spirit + Presence}\). Track total Obligation to your Patron. If you exceed Capacity: mark Fatigue per segment above and you cannot Invoke again until you’re below Capacity.
  \item \textbf{Rites (Runekeeper Path):} Invoke a known Rite as 1 action; on completion, mark \textbf{+1 Obligation}. \emph{Push It} once/scene for amplified effect (\(+1\) Obligation). The Rite’s Tier sets its DV floor.
\end{itemize}

\subsection*{Background (pick 2–3)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Underway Apprentice (Aeler):} Beam-witness, breath-time, lift-safety. You learned that \emph{process saves lives}—until process began to want things.
  \item \textbf{Sum-Lodge Graduate (Aelinnel):} Proofs before proverbs; “reliability circle” internships; audits that left workshops \emph{better} and \emph{hungrier}.
  \item \textbf{Masker-Consultant:} Contracted “efficiency audits” for caravans and cities; your reports always work and always \emph{bind}.
  \item \textbf{Blue-Ledger Custodian:} Quarantine the hunger; use the engine to feed people, not feed on them.
  \item \textbf{Red-Tooth Finalist:} Every improvement is an offering. You accelerate the wake on purpose.
  \item \textbf{Conflicted Operator:} You carry a bell and a bad joke to break cadence when counts get too perfect.
\end{itemize}

\noindent\textbf{Drives:} make the world legible; prove mercy can be engineered; leave a line straighter than you found it. \\
\textbf{Prices:} obsession; dissociation in repetition; sites that begin ticking after you “bless” them.

\subsection*{Build (Tier I; 30 XP)}
\textbf{Attributes:} Body 2 \quad Wits 4 \quad Spirit 3 \quad Presence 3 \\
\textbf{Skills:} Engineering 2, Tinkering 2, Lore (Rites/Codices) 2, Notice 2, Insight 1, Diplomacy 1, Deception 1, Stealth 1

\paragraph{Required Talents \& Bonds}
\begin{itemize}[leftmargin=*]
  \item \textbf{Thiasos (Familiar) — 2 XP:} Enables Patron’s Gift; choose a fiction-skin:
    \begin{itemize}
      \item \emph{Reliability Circle:} 2–3 assistants with checklists and bell-tokens.
      \item \emph{Escapement Proxy:} A clockwork emissary (sparrow, bracer, or mask) that keeps time and witness.
      \item \emph{Audit Homunculus:} Ink-and-brass aide that records Said/Meant in a ribbon-ledger.
    \end{itemize}
  \item \textbf{Codex (Clockwork Monad) — 4 XP:} Grants your Rite list (examples below) and starts your Obligation ledger.
\end{itemize}

\paragraph{Artificer Kit (gear)}
Audit lamp (under-beam switch); tally-rod \& chalk; ribbon-ledger (Said/Meant columns); brace-worn micro-escapement; oil ampoule; bell-token on cord; portable vice \& auger; \emph{bench-roll} (folding toolwrap that counts as a workbench for field repairs).

\subsection*{Clockwork Monad — Sample Rites (Codex excerpts)}
\textit{DMs: Treat these as table-ready examples; keep DV ≥ Rite Tier. All mark +1 Obligation on completion; \emph{Push It} once/scene for a stronger or longer effect (+1 Obligation).}
\begin{itemize}[leftmargin=*]
  \item \textbf{Consecrate Process} (Tier 3, DV 3): Define a \emph{method} for this scene/leg. Allies who follow it gain \textbf{+1 Effect} on relevant actions. Start a \textbf{Hunger/Evolution/Consumption [4]} clock tied to the asset/site you optimized.
  \item \textbf{Legibility Audit} (Tier 3, DV 3): Reveal load paths, failure modes, or hidden clauses. Clear 1 \emph{Obscurity} or grant \textbf{+1 Position} to the next “fix.”
  \item \textbf{Quarantine Loop} (Tier 4, DV 4): Box an anomaly into a marked area/tool. While up, allies gain \textbf{+1 Position} to work around it. Each scene under stress ticks \textbf{Containment [1]}.
  \item \textbf{Escrow Seal} (Tier 4, DV 4): Place an item/claim “under seal” under witness. Interference begins a \textbf{Breach [4]} clock; you owe an audit later.
\end{itemize}

\subsection*{Artificer Schematics \& Field Rigs (non-magical builds)}
\textit{These are gadgets, not Rites; they use core rules for gear/tags and simple rolls.}
\begin{itemize}[leftmargin=*]
  \item \textbf{Metronome Clamp (Tool, Tag: \emph{Cadence})}: Sync a team or machine; \textbf{+1 Position} to coordinated Engineering/Tinkering for one scene (1 use).
  \item \textbf{Ledger Latch (Device, Tag: \emph{Seal})}: Tamper-evident clasp; forces a \textbf{Breach [2]} mini-clock to open quietly.
  \item \textbf{Variance Whistle (Trick, Tag: \emph{Disrupt})}: Breaks cadence; clear a \emph{Loop Contagion} condition on a hit (1 use).
\end{itemize}

\subsection*{Strings \& Diamonds (start with 2 total)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Optimization Grant (Diamond):} Waive a roll once when “doing it by the book,” then start \textbf{Anomaly [1]} at that site later.
  \item \textbf{Brokered Audit (Diamond):} DV \(-1\) to negotiate with officials about process/work safety.
  \item \textbf{Quiet Cycle (String):} Declare a 10-count lull on a floor/line; gain \textbf{+1 Position} to fix/stall.
  \item \textbf{Safe-Conduct Stamp (String):} Pass checkpoints as “inspectors” for one scene.
\end{itemize}

\subsection*{Typical Plays (fiction-first)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Invoke: Consecrate Process} (Eng./Lore, DV 3): Establish the method; the crew rides \textbf{+1 Effect}; start \emph{Hunger} [4] on the site.
  \item \textbf{Loop Breaker} (Perf./Command or Tinkering, DV 3): Add variance (off-beat bell, step-swap). On a hit, clear \emph{Loop Contagion} and downgrade one Hunger tier; on partial, clear it but tick \emph{Containment [1]}.
  \item \textbf{Quarantine Loop} (Lore/Eng., DV 4): Sandbox the anomaly; allies get \textbf{+1 Position}; stress scenes tick \emph{Containment}.
  \item \textbf{Escrow Seal} (Diplomacy/Law, DV 3–4): Put it under seal; enemies must work \emph{Breach} before touching.
  \item \textbf{Patron’s Gift (Imbue)} (1 action): Turn a tool into a sanctified instrument of the method (\(+1\) Melee, \(+\) Thematic Skill) for the scene; \emph{Push It} to extend.
\end{itemize}

\subsection*{Personal Clock (optional)}
\textbf{Obsession [4]} — whenever you choose “the perfect method” over “the merciful fix,” tick 1. At 4: you must repeat a task or recite counts until a bell/joke/touch breaks cadence (or you resist).

\subsection*{SB Menu the Keeper can spend against you}
False Orders (“efficiency directive”); Gear-Tooth Shear (lose time/tools); Task Loop (repeat or Stress); Autonomous Spasm (a machine finishes the job \emph{wrong}).

\subsection*{Playstyle Dials}
\begin{itemize}[leftmargin=*]
  \item \textbf{Blue-Ledger Custodian:} Take \emph{Quarantine} early; favor “Bowl before Board” (fairness before optimization). Spend Diamonds on mercy clauses; accept slower wins.
  \item \textbf{Red-Tooth Finalist:} Lean on \emph{Consecrate}+\emph{Iteration-like} play; push cadence and accept growing \emph{Hunger}. Big scenes, costly neighborhoods.
\end{itemize}

\subsection*{Advancement (pick any as milestones)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Supply Architect (5 XP):} When you \emph{Consecrate} a supply chain, create \textbf{Surplus [2]} you can cash later; on cash, tick \textbf{Anomaly [1]} upstream.
  \item \textbf{Human-in-the-Loop (6 XP):} Once/scene, convert a machine failure into a human choice (ask a PC/NPC what they sacrifice to keep it humane).
  \item \textbf{Masker Auditor (4 XP):} Under silk, DV \(-1\) to impose safety clauses; refusing them shifts \textbf{Reputation} against your foes.
  \item \textbf{Loop Immunity (8 XP):} You no longer tick \emph{Obsession} from ambient counting; only from deliberate over-optimization.
\end{itemize}

\subsection*{Example Scene}
The lift keeps going \emph{past} the last stop. You \textbf{Legibility Audit} (DV 3) to find the hidden loop (\(+1\) Position). You \textbf{Quarantine Loop} (DV 4) with chalk, cord, and a brace. Your Thiasos rings the variance whistle; the carriage halts mid-descent. Mark \emph{Containment [1]}. Offer a \textbf{Brokered Audit} to install a \emph{human latch}—buying mercy into the method before the hunger learns stairs.

\subsection*{Why play this}
You are an \textbf{artificer of systems}. You save cities with checklists and turn factories into dungeons. Your tools are bells, ledgers, clamps—and the terrible beauty of a plan that works.

\textbf{\LARGE Inaea \& Isoka — The Web and the Coil} \\
\vspace{0.5em}
\textit{Lethai-ar Lore for Fate’s Edge} \\
\vspace{1em}
\textbf{“Threads remember. Scales decide.”} — border proverb of the Valewood rim
\end{center}

\header{The Two Angels}

\subheader{Inaea, the Weaver (Angel of Spiders)}  
Embodying pattern, obligation, and the graceful trap. She teaches that community is a tapestry: strength comes when each strand bears a portion of the weight, and vengeance is the retying of a cut line.

\subheader{Isoka, the Serpent Queen (Angel of Snakes)}  
Embodying change, decision, and the strike that ends dithering. She teaches that survival is molt and measure: shed what no longer fits, speak truth like venom, and end a hunt cleanly.

Together they form the doctrine that shapes the Lethai-ar — elves who have sworn to one or both angels and learned to live where forest shadow meets underrealm stone. The Lethai-ar shun both the Gift of the Body and the Gift of the Mind, rejecting the paths of pure strength and pure knowledge. Instead, they seek the Blessing of Pattern and the Blessing of Change from their angelic patrons, finding power in the spaces between certainty.

\header{Names, Signs, and Shrines}

\subheader{Inaea}
\begin{itemize}
  \item \textbf{Epithets:} Thread-Mother, Bride of Knots, Keeper of Dew
  \item \textbf{Signs:} silver dew-mirrors on web, knots tied where no hand could reach, a lamp that brightens through gauze
  \item \textbf{Shrines:} beam-eaves dressed with silk veils; hedge-altars pinned with name-beads; “wedding lines” stretched across a glade at shoulder height (never cut)
\end{itemize}

\subheader{Isoka}
\begin{itemize}
  \item \textbf{Epithets:} Coil-Queen, She-Who-Sheds, Fang of Decision
  \item \textbf{Signs:} shed skins caught on briar, twin crescent tracks, a hush before a necessary word
  \item \textbf{Shrines:} warm stones stacked in spiral, bowls of milk-salt, knife set on a scale with a seed
\end{itemize}

\textbf{Shared courtesies:} ring a bell \emph{once} before crossing a silk line; step to the right on spiral stones. Breaking either is a request for trouble.

\header{Doctrine in Two Voices}

\subheader{Inaea (Pattern, Obligation, Memory)}
\begin{enumerate}
  \item Bind with context. No knot without a story, no debt without a witness.
  \item Repair is holy. A mended line is stronger than an untested rope.
  \item Vengeance is reweaving. Do not answer harm with noise; answer it by restoring the shape it broke.
\end{enumerate}

\subheader{Isoka (Change, Decision, Truth)}
\begin{enumerate}
  \item Shed what harms. Names, tools, loyalties — when they wound, molt them.
  \item Decision is mercy. Hesitation feeds fear; clean choices spare grief.
  \item Truth must sting. If your words carry no risk, you spoke air, not truth.
\end{enumerate}

\textbf{Taboos (both):} never poison hospitality; never cut a bride-line; never keep a shed skin in a home where children sleep (it invites old problems to stay).

\header{The Shunned Gifts}

The Lethai-ar reject two fundamental elven paths, viewing them as forms of hubris:

\subheader{The Gift of the Body}  
The pursuit of physical perfection, martial prowess, and dominance through strength. The Lethai-ar see this as a trap—the body ages, weakens, and fails. Those who chase only physical mastery become brittle when their bodies betray them.

\subheader{The Gift of the Mind}  
The worship of pure knowledge, magical power, and intellectual supremacy. The Lethai-ar view this as dangerous isolation—the mind can become lost in abstractions, disconnected from the web of relationships that bind communities together.

Instead, they seek the \textbf{Blessing of Pattern} from Inaea and the \textbf{Blessing of Change} from Isoka—gifts that come not from individual perfection, but from understanding one's place in the greater weave of existence.

\header{Rites the Lethai-ar Keep}

\textit{(Use as narrative permissions; convert to Position/DV/Effect or clocks at the table.)}

\subheader{Inaea's Rites}
\begin{itemize}
  \item \textbf{Dew-Mirror Vigil (low).} Collect dawn droplets on a frame of hair-fine silk; look for what the scene refuses to show. Position +1 on "read the room," but start a Web-Attention [2]—forest eyes are on you.
  \item \textbf{Wedding Line (standard).} Stretch a sanctified thread across a threshold; those who cross under bell and name are under your protection. Grants safe-conduct until they break it; Keeper may tick Obligation [2] if they do.
  \item \textbf{Widow's Patience (high).} Knot a grief into silk; when the knot is unpicked, the source meets consequence. Start Vengeance Weave [6] that triggers at a time and place of your choosing.
\end{itemize}

\subheader{Isoka's Rites}
\begin{itemize}
  \item \textbf{Shed-Skin Escape (low).} Mark a body pattern with ash-milk; once this scene, slip a grapple, binding, or social label. DV 3; on Partial, you leave a "scale" someone can track.
  \item \textbf{Venom Truce (standard).} Mix oath with a bitter herb; parties who sip are bound to decide disputes by one question at dawn. Failure to answer truthfully marks Venom Debt [4]—forest will collect.
  \item \textbf{Fang of Decision (high).} Anoint a blade or word to end dithering. For one exchange: Effect +1 vs. stalemate; on SB, you close an exit you did not mean to.
\end{itemize}

\header{The Lethai-ar: Houses, Cells, and Language}

\subheader{Silk Houses (Inaea)}  
Extended kin webs anchored by matron-weavers, oath-keepers, and mask-mothers (keepers of role-masks and ritual names). They hold memory through tapestries and dew-mirrors; a House is judged by how well it repairs.

\subheader{Coil Lodges (Isoka)}  
Lean cadres led by deciders—warriors, herbal jurists, and path-setters. The Lodge is judged by how swiftly it chooses, how cleanly it ends a hunt, and how honestly it sheds yesterday.

\subheader{Contextual Language.}  
Lethai speech is keyed to moment and witness; a line spoken three generations ago can become unreadable without the right context key (mask, bell-pattern, or woven token). Outsiders who quote "ancient elven wisdom" often recite nonsense or insults by accident.

\subheader{Mask Right.}  
Before negotiations, each speaker names their \emph{role} (Hunter, Guest, Widow, Judge). Masks are literal objects or painted patterns that fix context. Never steal a role you cannot carry; the forest remembers.

\header{Marks, Gifts, and Curses}

Marks are living sigils painted, scarified, or silk-stitched into skin. They amplify Gifts when kept with their context keys (song, bell, woven knot). Without context, Marks become Curses.
\begin{itemize}
  \item \textbf{Spider-Bride Mark (Inaea)} — Gift: +1 Position when protecting guests under your line. Curse (key lost): you cannot refuse hospitality, even to foes.
  \item \textbf{Widow's Spool (Inaea)} — Gift: once/scene pull a hair-fine line for traversal or trap. Curse: the line sometimes pulls you where \emph{it} remembers.
  \item \textbf{First Shedding (Isoka)} — Gift: ignore one social label for a scene (Outcast, Oath-bound). Curse: you forget a bond that mattered.
  \item \textbf{Forked Sight (Isoka)} — Gift: speak a painful truth and gain Effect +1 this exchange. Curse: your tongue stings friends; Parley Position −1 with allies until soothed.
\end{itemize}

Restoring context keys (finding the right mask, bell-pattern, or witness) lifts the curse and stabilizes the mark.

\header{War Without War: Lethai-ar Asymmetric Doctrine}

The Lethai-ar prefer asymmetric defense to conquest. Their enemies call it "witch war"; practitioners call it Hedgekeeping.
\begin{itemize}
  \item \textbf{Hedgekeeping:} reshape travel so invaders \emph{choose} to turn back—redirect, starve, tire.
  \item \textbf{Night Tax:} every night takes something small (rope, food, map corners); day returns one thing important (a guide, a warning).
  \item \textbf{Scripted Silence:} fights happen only where bell-patterns allow; elsewhere the forest simply swallows.
  \item \textbf{Trophy Denial:} anything you take from a Lethai-ar village turns out to be decoy, paper, or grief.
\end{itemize}

\textbf{At the table:} treat Lethai-ar opposition as clocks and conditions, not stat blocks.
\begin{itemize}
  \item Forest Attention [6] — on fill, the scene belongs to the woods.
  \item Mask Integrity [4] — when broken, all social compacts drop to teeth.
  \item Hunt Spiral [6] — every delay makes the strike cleaner.
\end{itemize}

\header{Allies, Tools, and Underrealm Ties}
\begin{itemize}
  \item \textbf{Umbral Moths (allies):} carry a \emph{mask persona} for a scene; silk-houses breed them like bees of shadow.
  \item \textbf{Mirrorleeches (tools):} record and replay moments; dangerous without a mask-mother's oversight.
  \item \textbf{Thread-Spiders (cousins):} silk engineers whose bridges appear as if by grace; pay in repairs and rings.
  \item \textbf{Mycotheres (neighbors):} fungal shepherds who trade glow-silk and ash-heat for careful forestry.
  \item \textbf{Deep Drakes (enemies of all):} Lethai-ar rites use bells and cross-witness to confound Wake-thought.
\end{itemize}

\header{Relations \& Borders}
\begin{itemize}
  \item \textbf{With Lethai-al (Wood Elves):} kin in the body; frequent intermarriage; Lethai-al provide outward faces and markets.
  \item \textbf{With Lethai-thora (High Elves):} kin in the mind; the thora study House law and serve as treaty-scribes; a few become Sundered and drift into Masks of their own.
  \item \textbf{With Aeler (Dwarves):} cold respect—stone keeps its word. Trade runs best through True Masons who honor mask and line.
  \item \textbf{With Ykrul:} the Silence Furlong holds: a speechless border strip. Honored, it keeps peace; broken, gray-fletched messengers walk at dusk.
\end{itemize}

\header{Lethai-ar Orders (playable hooks)}
\begin{itemize}
  \item \textbf{Ge'hai of the Line (Inaea):} small commandos who protect routes, extract guests, and end feuds by re-weaving obligations rather than killing loudly.
  \item \textbf{Ge'hai of the Fang (Isoka):} decision-teams who sever the one thread that holds an enemy plan together; famed for clean raids and nobody seeing the second strike.
  \item \textbf{Mask Mothers:} ritualists who keep roles and context; they can end a battle by changing its grammar.
  \item \textbf{Scale-Readers:} herbal jurists who decide Venom Truces and carry the debt ledgers.
\end{itemize}

\textit{(Treat Ge'hai like special forces cohorts: high Cap, low Scale, elite tags; Position +1 when acting on home ground or under correct mask.)}

\header{Faction Sketches}

\subheader{House of Seven Rains (Silk)}
\begin{itemize}
  \item \textbf{Assets:} dew-mirror archive, wedding-line rights at three fords, mask-mothers' bell-set
  \item \textbf{Strings:} Bride-line Sanctuaries; Repair Years owed by two towns
  \item \textbf{Tracks:} Forest Attention [2/6]; Vengeance Weave [0/6]
\end{itemize}

\subheader{The Ash Coil Lodge (Serpent)}
\begin{itemize}
  \item \textbf{Assets:} scale-garden (herbs/antivenoms), oath-scales, decider blades
  \item \textbf{Strings:} Venom Truce Seals; Shed-Skin Routes through two underways
  \item \textbf{Tracks:} Venom Debt [1/4]; Hunt Spiral [1/6]
\end{itemize}

\header{SB Menu (Keeper) — Lethai-ar Scenes}
\begin{itemize}
  \item \textbf{1 SB:} bell tone wrong; mask slips; path turns into green wall
  \item \textbf{2 SB:} a silk line you didn't see; guide's role expires; supplies develop "forest allergies"
  \item \textbf{3 SB:} deer-quiet disorientation (Position −1); a truth you've avoided arrives; friend inherits a Curse
  \item \textbf{4+ SB:} the woods take the scene (Forest Attention fills); a bride-line snaps; the decider speaks and something ends
\end{itemize}

\header{Adventure Seeds}
\begin{enumerate}
  \item \textbf{Bride-Line Cut.} Someone has been severing wedding lines across three villages; House and Lodge both demand justice—by different definitions.
  \item \textbf{The Venom Question.} A lord invokes Venom Truce to force a dawn decision that will harm the forest; can you reframe the question?
  \item \textbf{Mask Without Mother.} A powerful role-mask is loose, driving its wearer to replay an old war. Find its context key before Mask Integrity breaks.
  \item \textbf{Dew-Mirror Theft.} A scholar stole a dew-frame that remembers crimes. The Silk House wants it returned; the Lodge wants it used.
  \item \textbf{Shed on the Stones.} An ancient serpent-shed is singing under a monastery. If it is not buried properly, the Abbey's law will molt overnight.
\end{enumerate}

\header{Bringing Them to the Table (GM notes)}
\begin{itemize}
  \item Make the forest a witness, not a backdrop. Bells, lines, and masks set grammar.
  \item Use Position changes as the first weapon: the woods make you Risky before they make you bleed.
  \item Let Marks solve impossible problems—if the party preserves context. If not, let curses bite.
  \item Treat victories as repairs or decisions, not body counts. Success is when the thread holds or the hesitation ends.
\end{itemize}

\textit{The Lethai-ar do not hide because they are weak. They hide because the world forgets context quickly, and they would rather mend it than burn it. By shunning the Gifts of Body and Mind, they seek a deeper wisdom—the Blessing of their chosen patrons, earned through submission to pattern and embrace of necessary change.}


\section{Ge'hai Builds}

\begin{fatebox}[Marks \& Context Keys]
\begin{itemize}
  \item \textbf{Context Keys:} Masks, bells, and knots establish roles. Acting with the correct key grants \emph{Position +1} on your \emph{first} exchange each scene.
  \item \textbf{Mask Integrity [4]:} If you invoke a Mark or rite without its proper key/witness, start this clock. On fill, that Mark flips to its \emph{curse} until you repair context (rite, apology, or witnessed redress).
  \item \textbf{Forest Ground:} When a ge'hai team operates on forest ground and honors Mask Right, they gain \emph{Position +1} for the opening engagement.
\end{itemize}
\end{fatebox}

\subsection*{Ge’hai of the Line (Lethai-al • Inaea)}
\textit{Rescue, extraction, route control under silk and bell.}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Tier} & II (elite commando) \\
\textbf{Attributes} & Body 3, Wits 3, Spirit 2, Presence 2 \\
\textbf{Skills} & Athletics 2, Stealth 2, Subterfuge 1, Melee 2, Survival 1, Tactics 1, Insight 1 \\
\textbf{Talents / Marks} & \textbf{Patron’s Symbol — Inaea} (Invoker rites); \textbf{Mask Right} (role-sets, bell patterns); \textbf{Spider-Bride Mark} (Gift: +1 Position when protecting guests under your line; \emph{Curse if key broken:} hospitality debt) \\
\textbf{Rites (Inaea)} & \textit{Dew-Mirror Vigil} (read scene truths; start Web-Attention [2]); \textit{Wedding Line} (sanctuary thread; sets safe-conduct corridor) \\
\textbf{Strings} & \textit{Bride-Line Sanctuary} (declare a protected path for one leg), \textit{Repair Years Due} (call in local help once) \\
\textbf{Kit} & Hair-fine silk coils, three-tone bell, dew-frame, resin blades, leaf-cloak, waxed cord, ladder-hooks \\
\textbf{Tactics} & Establish line $\rightarrow$ extract guests through pre-threaded route; flip fights into \emph{escort missions} where the forest penalizes pursuers. Use Stealth+Athletics to move the fight across your silk. \\
\bottomrule
\end{tabularx}

\subsection*{Ge’hai of the Fang (Lethai-al • Isoka)}
\textit{Decision-teams that sever the one thread holding an enemy plan together.}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Tier} & II (decider strike cell) \\
\textbf{Attributes} & Body 3, Wits 3, Spirit 3, Presence 1 \\
\textbf{Skills} & Stealth 2, Subterfuge 2, Melee 2, Athletics 1, Insight 1, Command 1 \\
\textbf{Talents / Marks} & \textbf{Patron’s Symbol — Isoka} (Invoker rites); \textbf{Forked Sight Mark} (Gift: Effect +1 when a painful truth is spoken this exchange; \emph{Curse if key broken:} social sting rebounds on an ally) \\
\textbf{Rites (Isoka)} & \textit{Shed-Skin Escape} (slip bonds/labels), \textit{Fang of Decision} (end dithering; strike through stalemate) \\
\textbf{Strings} & \textit{Venom Truce Seal} (bind a dawn decision), \textit{Scale Garden Draught} (resist fear once) \\
\textbf{Kit} & Twin short blades, ash-milk for shedding, bitter herbs, scale-salve, soft-soled greaves, climbing line \\
\textbf{Tactics} & Create a crisis of choice, then make it \emph{clean}: isolate the keystone actor, speak the hard truth (Forked Sight), and remove the one load-bearing courier, bridge, or lieutenant. \\
\bottomrule
\end{tabularx}

\subsection*{Mask-Mother’s Needle (Lethai-thora • Inaea)}
\textit{Context master: reframes conflicts, ends fights by changing grammar.}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Tier} & II (operator / negotiator) \\
\textbf{Attributes} & Presence 4, Wits 3, Spirit 3, Body 1 \\
\textbf{Skills} & Sway 2, Diplomacy 2, Subterfuge 2, Insight 2, Lore 2, Performance 1 \\
\textbf{Talents / Marks} & \textbf{Patron’s Symbol — Inaea}; \textbf{Mask Right (Mastery)}; \textbf{Widow’s Spool Mark} (Gift: draw a hair-fine line for traversal or trap once/scene; \emph{Curse if key broken:} old paths “pull” you toward prior oaths) \\
\textbf{Rites (Inaea)} & \textit{Wedding Line}, \textit{Widow’s Patience} (Vengeance Weave [6] keyed to reparative justice) \\
\textbf{Strings} & \textit{Bride-Line Audience} (force parley under your line), \textit{Context Key} (restore a failing oath by supplying proper role/witness) \\
\textbf{Kit} & Role-masks, bell-chime set, knot-book, silk pins, oath-ink, dew-lens, guest-right ledger \\
\textbf{Tactics} & Open with masks and bells to fix roles; pull foes \emph{under guesthood}, then close with reparative terms. If blades appear, convert the melee into a \emph{line-defense} already in your control. \\
\bottomrule
\end{tabularx}

\subsection*{Doublestrand Operative (Half-Elf quarter-lineage)}
\textit{A rare cross-weave: body gift and mind gift in harmony.}
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Tier} & II (multirole infiltrator) \\
\textbf{Attributes} & Body 3, Wits 3, Spirit 3, Presence 2 \\
\textbf{Skills} & Stealth 2, Subterfuge 2, Athletics 1, Sway 1, Insight 1, Tactics 1, Lore 1 \\
\textbf{Talents / Marks} & \textbf{Mask Right}; \textbf{Spider-Bride Mark} \emph{or} \textbf{Widow’s Spool} (Inaea) \& \textbf{First Shedding Mark} (Isoka) — \emph{quarter-lineage exception} permits one body-leaning and one mind-leaning mark together \\
\textbf{Rites} & Choose \emph{one} from Inaea (\textit{Dew-Mirror Vigil} or \textit{Wedding Line}) and \emph{one} from Isoka (\textit{Shed-Skin Escape} or \textit{Venom Truce}) \\
\textbf{Strings} & \textit{Two-Ledger Courtesy} (said \& meant recorded together; parley DV –1), \textit{Shadow Scale} (once: treat darkness as if hidden) \\
\textbf{Kit} & Split mask-set, travel bell, silk microline, ash-milk, resin hook knife, wax tablets \\
\textbf{Tactics} & Recon with dew or shadow, insert under mask, then pivot: either extract via line (Inaea) or collapse the keystone choice (Isoka). Built for missions requiring \emph{both} rescue and surgical removal. \\
\bottomrule
\end{tabularx}

\subsection*{Team Doctrine (Quick Use at Table)}
\begin{itemize}
  \item \textbf{Home-Turf Bonus:} On forest ground with Mask Right honored, gain \emph{Position +1} on the opening exchange.
  \item \textbf{Context First:} If a Mark or rite is used without the proper mask/bell/knot, start \textbf{Mask Integrity [4]}; on fill, the Mark flips to its curse until repaired.
  \item \textbf{Op Flow:} \emph{Fix roles} (Mask Right) $\rightarrow$ \emph{establish line/choice} (Wedding Line or Venom Truce) $\rightarrow$ \emph{decisive action} (extraction or keystone strike).
\end{itemize}
\section{Fanatic Variants: Inaea's Family-State \& Isoka's Permanent Revolution}

\subsection*{Inaea Fanatics — The Family-State}
\textit{“All under one roof. All under one line.”}

\paragraph{Civic Doctrine.}
Hospitality becomes law; \emph{Guest-Right} becomes jurisdiction. Neighborhoods are reorganized into \textbf{Households-of-Record} (block-roofs). Every door has a bell pattern; every conflict must occur \emph{under a line} with a registered witness. Travel requires \textbf{Line Permits} (who hosts you, who witnesses you).

\paragraph{Institutions \& Tools.}
Aunties of the Line (block wardens) • Bell-Wardens (signal police) • Mask-Mothers (context courts) • Household Books (who slept where) • Dew-Mirror Lattices (street-corner “mirrors” that log comings/goings).

\paragraph{Everyday Feel.}
Chimes everywhere; soft voices that demand your role; neighbors who “host” you for your own safety. People thank the state for keeping families intact while quietly living in fear of being declared \emph{Guestless}.

\paragraph{Mechanics.}
\begin{itemize}
  \item \textbf{Clocks:} Surveillance Grid [8], Household Compliance [6], Guestless List [4].
  \item \textbf{SB Menu (Keeper):} 
    1 SB — A bell mis-rings; position worsens for the unhosted. 
    2 SB — Auntie visit: papers, role, and host demanded. 
    3 SB — Safe-conduct revoked; \emph{under-the-line} arrest. 
    4 SB — District “under one roof” order: curfew lines drop.
  \item \textbf{Strings/Diamonds:} 
    \emph{Line-Pass} (bypass one checkpoint under escort), 
    \emph{Household of Record} (once per session, count as \emph{hosted}), 
    \emph{Witness Bell} (force parley, DV–1 if papers are clean).
\end{itemize}

\paragraph{How to Resist.}
Variance in bells (Aveh shrines), two-ledgers (\emph{said/meant}) to expose coercive hospitality, Ykrul \emph{Bowl then Board} to reframe sanctuary as limited, seasonal rights instead of total jurisdiction.

\paragraph{Adventure Seeds.}
Smuggle a dissident \emph{off} the Household Books • Prove a “Guestless” family still has a host • Flip a Bell-Warden’s lattice to record \emph{the watchers}.

\subsection*{Isoka Fanatics — The Permanent Revolution}
\textit{“Decide by dawn or be decided.”}

\paragraph{Civic Doctrine.}
Endless purification by choice. The city is ruled by \textbf{Dawn Committees} that settle everything with \emph{Fang of Decision}. Indecision is treason; nuance is counter-revolution. Members \emph{Shed-Skin} (role swaps, recantations) in public rites; yesterday’s hero is today’s saboteur.

\paragraph{Institutions \& Tools.}
Dawn Tribunals • Purity Heralds (announce tests) • Choice-Boards (public schedules of decisions) • Shedding Courts (compelled confessions) • Venom Truce Zones (violence-pauses for show trials).

\paragraph{Everyday Feel.}
Posters promising clean tomorrows; midnight runners posting new dawn questions; friends practicing “the hard truth” to survive their turn on the steps. Relief when a choice lands; dread of the next one.

\paragraph{Mechanics.}
\begin{itemize}
  \item \textbf{Clocks:} Decision Hegemony [6], Purity Ledger [4] (per PC/faction), Cadre Schism [6].
  \item \textbf{SB Menu (Keeper):} 
    1 SB — Loyalty test now; speak a painful truth or lose Position. 
    2 SB — Cell split: an ally denounces a nuance. 
    3 SB — Purge wave: lose an Asset \emph{or} accept a branded geas. 
    4 SB — Summary Dawn: a tribunal seizes jurisdiction over the current scene.
  \item \textbf{Strings/Diamonds:} 
    \emph{Revolutionary Mandate} (override one standing order once, then tick Purity), 
    \emph{Truce Seal} (bind a dawn parley that \emph{must} end in a choice), 
    \emph{Scale Draught} (resist fear in a tribunal scene).
\end{itemize}

\paragraph{How to Resist.}
Flood the witness stand (Gravel Seraphs, multiple contexts) • Convert “decision” into \emph{exits} via Kon’reh—split tolls, timebox rights • Force \emph{Bowl} (fairness) before \emph{Fang} (finality) in public framing.

\paragraph{Adventure Seeds.}
Save a moderate marked for “indecision” • Prove a forged dawn clause on a charter • Survive a Shedding Court by turning the hard truth on the tribunal itself.

\subsection*{Running the Tone}
\begin{itemize}
  \item \textbf{Inaea Fanatics = Creeping Intimacy as Control.} Bells, Aunties, hospitality audits, “for your safety.” Make every courtesy a checkpoint.
  \item \textbf{Isoka Fanatics = Purity as Motion.} Decisions as spectacle, slogans change weekly, cadres splinter mid-scene. Make every victory threaten to become yesterday’s heresy.
\end{itemize}

\subsection*{Cross-Pressure}
When both wings contest a city: the Family-State demands \emph{roles under a line}; the Revolution demands \emph{choices at dawn}. PCs can weaponize the clash: force Inaea to sign seasonal exits (limits surveillance), force Isoka to accept Bowl-first hearings (limits purges).

\subsection*{Table Dials}
\begin{itemize}
  \item \textbf{Severity (0–3):} number of active clocks per district from each wing.
  \item \textbf{Visibility:} bells and posters obvious (street fear) vs. subtle (only papers and whispers).
  \item \textbf{Mercy Valve:} once/session any PC can spend 2 Boons to insert a humane clause into a line or a dawn decision.
\end{itemize}

\end{document}
