\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning}
\usepackage{mdframed}
\usepackage{sectsty}
\usepackage{needspace}
\usepackage{changepage}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{marginnote}

Ōēźometry{margin=1in}
\setstretch{1.2}
\sectionfont{┼éarge}
\subsectionfont{\normalsize}
Ō½Ģsectionfont{\small}

ŌēĀwmdenv[linecolor=black,linewidth=1pt,backgroundcolor=gray!5]{adventurebox}
ŌēĀwmdenv[linecolor=gray,linewidth=0.5pt,backgroundcolor=gray!3]{mechanicbox}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Sands of Moon & Brass}
\fancyhead[R]{Fate's Edge Module}
\fancyfoot[C]{├Šepage}

\title{Sands of Moon & Brass \\ ┼éarge Deserts of Galaninan, Fhara, and Ashaan (v0.1)}
\author{}
\date{}

\begin{document}

\maketitle

\begin{center}
\textbf{A cultureŌłÆfocused module for desert trade, nomad law, starŌłÆpath navigation, and qanat politics by the caravanserai folk of the Galaninan Great Erg, Fharan Marches, and Ashaani Provinces. Built to slot into Caravans, Wilderness, Sea, Political Intrigue, and Psionics using core Fate's Edge procedures (Position/DV, SB, timers, Strings, Favor/Leverage/Exposure).}
\end{center}

\section*{Design Goals}
\begin{itemize}
Ōä╣tem \textbf{Etiquette with teeth.} Water first, lampŌłÆlit contracts, and title respect shape play as much as blades.
Ōä╣tem \textbf{CaravanŌłÆŌłÆnomad parity.} Trade routes, tribal law, and oasis politics are equally supported.
Ōä╣tem \textbf{Seamless integration.} Uses existing market rules, faction timers, and currenciesŌłÆŌłÆŌłÆno new dice math.
Ōä╣tem \textbf{Low bookkeeping.} Caravan, tribe, season wheel, and 2ŌłÆŌłÆ3 visible timers per scenario.
\end{itemize}

\section*{Quickstart (2 minutes)}
\begin{enumerate}
Ōä╣tem Make a Caravan/Tribe Sheet (┬¦1) and a Mount/Camel Sheet (┬¦2). Pick 1ŌłÆŌłÆ2 Strings (waterŌłÆshare tablet, lampŌłÆoath seal, windŌłÆtower charter) and 3ŌłÆŌłÆ4 Tags.
Ōä╣tem Mark the Season Wheel (┬¦3) and current Theaters (erg, marches, provinces). Start Standing [6], Feud [4], Exposure [6].
Ōä╣tem Pick an opening Score Type (┬¦4): Desert Caravan, Oasis Siege, Qanat Sabotage, or Council Moot.
Ōä╣tem Frame the approach, set Position/DV from tags/venue, play. On any 1, GM spends from Sands & Brass SB (┬¦6.4).
\end{enumerate}

\section{Caravan/Tribe Sheet (Template)}

\begin{center}
\begin{longtable}{|p{5cm}|p{5cm}|}
ŌäÄline
\textbf{[CARAVAN / TRIBE NAME]} & \\
ŌäÄline
Seat (home oasis): ____________ & Council Affiliation: ____________ \\
ŌäÄline
\end{longtable}
\end{center}

\textbf{Strings (2ŌłÆŌłÆ3):} waterŌłÆshare tablet $ŌĆó$ lampŌłÆoath seal $ŌĆó$ trade charter $ŌĆó$ feud settlement $ŌĆó$ wayŌłÆsong $ŌĆó$ guideŌłÆstone

\textbf{Tags (2ŌłÆŌłÆ4):} WaterŌłÆWise $ŌĆó$ StarŌłÆPathed $ŌĆó$ HeatŌłÆTrained $ŌĆó$ MerchantŌłÆKnown $ŌĆó$ Hospitable $ŌĆó$ LawŌłÆStrict $ŌĆó$ PoetŌłÆLoud $ŌĆó$ FalconŌłÆBanner

\textbf{Tracks:}
\begin{itemize}
Ōä╣tem Standing [6] (reputation among desert folk and cities)
Ōä╣tem Feud [4] (active quarrel; on fill $ŌåÆ$ bloodŌłÆprice owed or war)
Ōä╣tem Exposure [6] (to foreign powers; for nonŌłÆdesert venues)
\end{itemize}

\textbf{Oath Ledger:} favors owed/held (name + what; acts like Favor/Leverage within desert spheres)

\textbf{Notables:} guide $ŌĆó$ warŌłÆleader $ŌĆó$ speaker $ŌĆó$ poet $ŌĆó$ waterŌłÆkeeper

\textbf{Standing [6]:} Tick up for kept oaths, fair trade, valor; tick down for oathŌłÆbreaking, cowardice, or sacrilege. High Standing grants Audience: Respectful at council; low invites Audience: Skeptical.

\textbf{Feud [4]:} Name the counterparty. Ticks on insult, theft, or harm; clear by bloodŌłÆsilver, ordeal, or deed.

\section{Mount/Camel Sheet (Dromedaries & Chargers)}

\textbf{Pick a mount and 3ŌłÆŌłÆ4 tags.}

\textbf{Dromedary (caravan):} waterŌłÆefficient, steady pace, cargo capacity.

\textbf{Charger (war):} fast sprint, agile turns, combatŌłÆtrained.

\textbf{Sand Runner (scout):} light, narrow feet, collapsible saddle for dunes.

\textbf{Mount Tags (choose 3ŌłÆŌłÆ4)}
\begin{itemize}
Ōä╣tem SandŌłÆSure ŌłÆŌłÆŌłÆ Desert sand Drive/Handle DV ŌłÆŌłÆ1; dune ridges, quicksand.
Ōä╣tem WaterŌłÆEfficient ŌłÆŌłÆŌłÆ Half water consumption; ignore first Fatigue from heat.
Ōä╣tem NightŌłÆRunner ŌłÆŌłÆŌłÆ In darkness, Navigate DV ŌłÆŌłÆ1; Ambush at Dawn starts one Position higher.
Ōä╣tem HeatŌłÆReady ŌłÆŌłÆŌłÆ In sun/fire scenes, first Condition +1 is ignored.
Ōä╣tem QanatŌłÆSwimmer ŌłÆŌłÆŌłÆ Underground water passages; Disengage gains Position +1 in tunnels.
Ōä╣tem FalconŌłÆEye ŌłÆŌłÆŌłÆ +1 die on Lookout/Watches; spot distant smoke/sand columns.
Ōä╣tem BrassŌłÆBearer ŌłÆŌłÆŌłÆ Carries heavy loads; ignore encumbrance penalties for trade goods.
Ōä╣tem Poet's Drum ŌłÆŌłÆŌłÆ Once/score gain Audience: Fierce aboard; can rally morale.
Ōä╣tem GuideŌłÆStones ŌłÆŌłÆŌłÆ Hidden cairnŌłÆmarks; once/score Navigate Position +1 in trackless sand.
\end{itemize}

\textbf{Roles (choose at table)}\\
Guide (navigation), WarŌłÆLeader (combat/assault), Speaker (parley), Poet (morale/saga), WaterŌłÆKeeper (supplies/repairs), Lookout (weather/sandstorm).

\section{Season Wheel & Theaters}

Spring Caravan $ŌåÆ$ Summer Heat $ŌåÆ$ Autumn Trade/Settling $ŌåÆ$ Winter Council.

At each transition:
\begin{itemize}
Ōä╣tem Advance Politics (Mandate/Crisis) for desert powers.
Ōä╣tem Roll Weather/Heat [4ŌłÆŌłÆ6] for the theater in play (erg, marches, provinces).
Ōä╣tem Offer Oath Opportunities (escort, feud settlement, shrine warding).
\end{itemize}

\textbf{Theaters}
\begin{itemize}
Ōä╣tem Galaninan Erg (central): vast dunes, oases, ancient ruins, starŌłÆtemples.
Ōä╣tem Fharan Marches (north): rocky plateaus, seasonal rivers, nomad camps.
Ōä╣tem Ashaani Provinces (south): qanat cities, date gardens, merchant houses.
\end{itemize}

\section{Score Types & Procedures}

Pick approach (Deceit $ŌĆó$ Speed $ŌĆó$ Shock $ŌĆó$ Parley) $ŌåÆ$ set Position/DV from venue/tags $ŌåÆ$ roll.

\subsection{Desert Caravan (Objective: Profit / Passage)}

\textbf{Timers:} Alarm [4ŌłÆŌłÆ6], Trade [6ŌłÆŌłÆ8], Mount Damage [4], BloodŌłÆPrice [4].

\textbf{Entry:} sand gate, hidden wadi, qanat shortcut, market ruse.

\textbf{On 1s:} GM spends SB $ŌåÆ$ sandstorm warning, hidden quicksand, rival caravan ambush, guide betrayal.

\textbf{Resolution:} When Trade fills, choose coin or Strings (trade charter, seasonal tithe, guide contract). If BloodŌłÆPrice fills, mark Feud +1 or pay bloodŌłÆsilver (Favor loss or obligation timer).

\subsection{Oasis Siege (Objective: Water / Tribute)}

\textbf{Timers:} Thirst [4], Defense [6], Sentries [4], Terms [6].

\textbf{Position tweaks:} WaterŌłÆEfficient/HeatŌłÆReady help; sandstorms/hail hurt.

\textbf{On 1s:} qanat gets blocked, arrows from palm grove, sandŌłÆburst collapses wall.

\subsection{Qanat Sabotage (Objective: Flood / Blockade)}

Use Market from Caravans/Wilderness; add desert perks:
\begin{itemize}
Ōä╣tem WaterŌłÆShare Tablets count as String once/session in desert venues.
Ōä╣tem Poet's Drum can convert Audience: Warm $ŌåÆ$ Favor (narrow) once/score with a verse.
\end{itemize}

\subsection{Council Moot (Objective: Law / Settlement)}

\textbf{Venue:} circle of stones, winter pavilion, qanat meeting hall.

\textbf{Moves:} OathŌłÆSwear (commit under penalty), Wager Wyrd (ordeal by feat), BloodŌłÆSilver (compensation roll), Witness the Verse (Poet stakes a truth).

\textbf{Outcomes:} resolve Feud, write Oath to ledger (acts as durable String), assign seasonal water rights or trade route.

\section{Etiquette, BloodŌłÆSilver, & Standing}
\begin{itemize}
Ōä╣tem \textbf{Etiquette (mechanics):} Water first grants Position +1; lampŌłÆlit contracts add +1 Effect; titles before names improve Audience tags.
Ōä╣tem \textbf{BloodŌłÆSilver (settlement):} Pay with coin, hostageŌłÆstring, or deed. Roll Petition/Broker vs DV 2ŌłÆŌłÆ4 (standing, witnesses, hurt). On hit, reduce Feud ŌłÆŌłÆ2; on partial, ŌłÆŌłÆ1; on miss, counterŌłÆoath demanded.
Ōä╣tem \textbf{Boasts & Verses:} A public boast creates Audience: Expectant; fulfill it to gain Standing +1, fail and mark Exposure or Feud.
\end{itemize}

\section{Desert & Qanat Procedures}

\subsection{DV Ladders}
\begin{itemize}
Ōä╣tem \textbf{Chase (Sand):} DV 2 dune ridge $ŌĆó$ 3 rocky pass $ŌĆó$ 4 salt flat $ŌĆó$ 5 sandstorm wall.
Ōä╣tem \textbf{Chase (Qanat):} DV 2 open tunnel $ŌĆó$ 3 narrow passage $ŌĆó$ 4 flooded section $ŌĆó$ 5 collapse zone.
Ōä╣tem \textbf{Board & Brace:} DV 3ŌłÆŌłÆ5 (tags: Charger helps lower DV).
\end{itemize}

\subsection{Weather/Heat Matrix [4ŌłÆŌłÆ6]}

Advance on 1s or fiction:
\begin{itemize}
Ōä╣tem \textbf{Khamsin/Sandstorm:} sight Position ŌłÆŌłÆ1; SandŌłÆSure cancels. On 1, Pursuit +1 (lost bearings).
Ōä╣tem \textbf{Flash Flood:} ranged actions ŌłÆŌłÆ1 die; on 1, Condition +1 (saddle/gear).
Ōä╣tem \textbf{Ward Storm:} Navigate DV +1; on 1, choose Delay (Distance stalls) or SandŌłÆBlast (Condition +1).
\end{itemize}

\subsection{Qanat Navigation & Hazards}

Treat as Cross Hazard (Body+Tactics/Craft) DV 3ŌłÆŌłÆ5; QanatŌłÆSwimmer/SandŌłÆSure grant Position +1. On 1, Tunnel Collapse or Water Breach.

\subsection{Sands & Brass SB (GM menu)}
\begin{itemize}
Ōä╣tem Hidden Quicksand: sudden sink; Mount Damage +1 unless Position was high.
Ōä╣tem Sandstorm Warning: dust wall rises; Pursuit +1 and Alarm +1.
Ōä╣tem Guide Betrayal: lose Position; find new path or fight lost.
Ōä╣tem Witness at Dune: a rival caravan sees; Exposure +1 (foreign) or Feud +1 (desert).
Ōä╣tem Verse Twisted: rumor flips an Audience tag against you.
\end{itemize}

\section{Desert Culture Tools (portrayal guidance)}
\begin{itemize}
Ōä╣tem Emphasize law and reciprocity (oaths, bloodŌłÆsilver, witness) over caricature.
Ōä╣tem Show plural livelihoods: traders, guides, waterŌłÆkeepers, poets, warriorsŌłÆŌłÆŌłÆnot only raiders.
Ōä╣tem Let women/elders hold seats and steer deals; avoid monolith tropes.
Ōä╣tem Lean into seasonality and council assemblies as civic life.
\end{itemize}

\textbf{Etiquette Hooks (once/scene in desert venues):} offer water first; name your grandfather's line; recite a verseŌłÆŌłÆŌłÆeach can grant Position +1 in parley.

\section{Factions & Fronts}
\begin{itemize}
Ōä╣tem \textbf{Desert Caravans (routes & dues):} StringsŌłÆŌłÆŌłÆpriority passage, guide contracts.
Ōä╣tem \textbf{Oasis Wardens:} StringsŌłÆŌłÆŌłÆwater rights, palm grove charters.
Ōä╣tem Ōä╣tem \textbf{Qanat Builders:} StringsŌłÆŌłÆŌłÆtunnel access, waterŌłÆshare tablets.
Ōä╣tem \textbf{StarŌłÆTemple Keepers:} StringsŌłÆŌłÆŌłÆsafeŌłÆrun chants, tithe.
Ōä╣tem \textbf{Night Moon Markets:} StringsŌłÆŌłÆŌłÆlampŌłÆoath seals, rare goods contracts.
\end{itemize}

\textbf{Front Timers (examples):}
\begin{itemize}
Ōä╣tem Rival Caravan Fleet [6ŌłÆŌłÆ8] (desert power organizes counterstrike).
Ōä╣tem Council Schism [6] (oath controversies split halls).
Ōä╣tem Qanat Drought [6] (trade mission needed; gain Standing if solved).
\end{itemize}

\section{Generators}

\subsection{Desert Targets (d66)}

11 starŌłÆtemple $ŌĆó$ 12 beacon tower $ŌĆó$ 13 cliff monastery $ŌĆó$ 14 salt works $ŌĆó$ 15 oasis fleet $ŌĆó$ 16 lord's pavilion $ŌĆó$ 21 dune storehouse $ŌĆó$ 22 rope walk $ŌĆó$ 23 ferryŌłÆchain $ŌĆó$ 24 dye shed $ŌĆó$ 25 customs tent $ŌĆó$ 26 river gate $ŌĆó$ 31 amber dunes $ŌĆó$ 32 quarry ridge $ŌĆó$ 33 ibex herd $ŌĆó$ 36camel yard $ŌĆó$ 35 sheep isle $ŌĆó$ 36 fortress gate $ŌĆó$ 41 smokehouse row $ŌĆó$ 42 eelŌłÆweirs $ŌĆó$ 43 guide stone $ŌĆó$ 44 winter oasis $ŌĆó$ 45 tax caravan $ŌĆó$ 46 river lighthouse $ŌĆó$ 51 shrine cove $ŌĆó$ 52 smugglers' cut $ŌĆó$ 53 patrol launch $ŌĆó$ 54 tollhouse $ŌĆó$ 55 grain pier $ŌĆó$ 56 sand wall breach $ŌĆó$ 61 ice slip $ŌĆó$ 62 wreckers' fires $ŌĆó$ 63 tide cave $ŌĆó$ 64 chain boom $ŌĆó$ 65 slate wharf $ŌĆó$ 66 mint caravan.

\subsection{Desert Hazards (d12)}

1 dune on a bend $ŌĆó$ 2 sudden flash flood $ŌĆó$ 3 quicksand pool $ŌĆó$ 4 deadwood snag $ŌĆó$ 5 scorpion nest $ŌĆó$ 6 heat shimmer $ŌĆó$ 7 hidden sideŌłÆcut $ŌĆó$ 8 rock echo $ŌĆó$ 9 toll chain halfŌłÆraised $ŌĆó$ 10 reeds conceal archers $ŌĆó$ 11 rainŌłÆswollen ford $ŌĆó$ 12 sandŌłÆsuck bank.

\subsection{Council Cases (d12)}

1 insult in verse $ŌĆó$ 2 stolen guide stone $ŌĆó$ 3 broken oath on winter grain $ŌĆó$ 4 bloodŌłÆprice disputed $ŌĆó$ 5 marriage claim $ŌĆó$ 6 wayŌłÆsong stolen $ŌĆó$ 7 hostage pledge lapsed $ŌĆó$ 8 salvage rights $ŌĆó$ 9 border cairn moved $ŌĆó$ 10 feud cooling terms $ŌĆó$ 11 verse witness contest $ŌĆó$ 12 mercenary pay withheld.

\subsection{Verses & Boons (d12)}

1 drumŌłÆsong that steadies hands $ŌĆó$ 2 falcon omen at dawn $ŌĆó$ 3 amber find $ŌĆó$ 4 guide's ghost shows a cut $ŌĆó$ 5 sandŌłÆstorm spares you $ŌĆó$ 6 omen of red sails $ŌĆó$ 7 poet's verse spreads $ŌĆó$ 8 winter pavilion adopts you $ŌĆó$ 9 qanat seal renewed $ŌĆó$ 10 cliff bell silent $ŌĆó$ 11 stars open path $ŌĆó$ 12 oathŌłÆring warms (truth told).

\section{Integration Notes}
\begin{itemize}
Ōä╣tem \textbf{Caravans:} Use trade market and caravan tools; desert Poet's Drum converts Audience to Favor once/score in market disputes.
Ōä╣tem \textbf{Wilderness:} Dune/overland jumps tie into outpost assets; desert hunts use Hunt/Chase ladders.
Ōä╣tem \textbf{Sea:} Swap staging at ports with oasis harbors; WaterŌłÆEfficient aids ship convoys in arid ports.
Ōä╣tem \textbf{Political Intrigue:} Treat the Council as a political venue; Standing sets default Position; BloodŌłÆSilver interacts with Favor/Exposure economies.
Ōä╣tem \textbf{Psionics:} Seers translate as omenŌłÆreaders; allow Psychic Weather Sense to soften sandstorm/heat penalties once/leg.
Ōä╣tem \textbf{Dragon's Lair:} Dunes and ancient ruins hide wyrm shrines; oaths may bind to ancient powers for perilous boons.
\end{itemize}

\section{Example of Play (short)}

\textbf{Setup:} Dromedary with SandŌłÆSure, NightŌłÆRunner, FalconŌłÆEye. Caravan Standing 3/6, Feud 1/4 (with Fharan Nomads). Score: Desert Caravan to Bir alŌłÆQamar (Trade).

\textbf{Approach:} ParleyŌłÆthenŌłÆSpeed. Speaker petitions for priority passage (DV 3). Partial $ŌåÆ$ Position stays Controlled, GM spends 1 SB.

\textbf{Action:} Guide triggers Navigate at night. NightŌłÆRunner drops DV to 2. Hit $ŌåÆ$ Trade +2; a 1 shows $ŌåÆ$ GM spends Sandstorm Warning (Alarm +1, Pursuit +1).

\textbf{Twist:} Flash Flood rolls in. SandŌłÆSure cancels Position penalty. WaterŌłÆKeeper finds a GuideŌłÆStone line: Navigate gains Position +1; Pursuit ŌłÆŌłÆ1.

\textbf{Close:} Trade fills. Caravan chooses a seasonal guide contract String instead of coin. BloodŌłÆPrice at 2/4; they pledge bloodŌłÆsilver at winter council to avoid Feud tick.

\section{GM Reference (one page)}
\begin{itemize}
Ōä╣tem \textbf{Caravan:} Standing [6] $ŌĆó$ Feud [4] $ŌĆó$ Exposure [6] $ŌĆó$ Oath Ledger.
Ōä╣tem \textbf{Mounts:} choose mount + 3ŌłÆŌłÆ4 tags. Roles: Guide $ŌĆó$ WarŌłÆLeader $ŌĆó$ Speaker $ŌĆó$ Poet $ŌĆó$ WaterŌłÆKeeper $ŌĆó$ Lookout.
Ōä╣tem \textbf{Season Wheel:} Spring caravan $ŌĆó$ Summer heat $ŌĆó$ Autumn trade/settle $ŌĆó$ Winter council.
Ōä╣tem \textbf{Score Types:} Desert Caravan $ŌĆó$ Oasis Siege $ŌĆó$ Qanat Sabotage $ŌĆó$ Council Moot.
Ōä╣tem \textbf{Key Timers:} Alarm $ŌĆó$ Trade $ŌĆó$ Mount Damage $ŌĆó$ BloodŌłÆPrice $ŌĆó$ Pursuit.
Ōä╣tem \textbf{Weather/Heat:} Khamsin/Sandstorm $ŌĆó$ Flash Flood $ŌĆó$ Ward Storm.
Ōä╣tem \textbf{SB Menu:} Hidden Quicksand $ŌĆó$ Sandstorm Warning $ŌĆó$ Guide Betrayal $ŌĆó$ Witness at Dune $ŌĆó$ Verse Twisted.
Ōä╣tem \textbf{Integration:} Dunes $Ōåö$ Wilderness $ŌĆó$ Market disputes $Ōåö$ Caravans $ŌĆó$ Council $Ōåö$ Political $ŌĆó$ Omens $Ōåö$ Psionics.
\end{itemize}

\section{Changelog}

\textbf{v0.1} ŌłÆŌłÆŌłÆ First draft: caravan/tribe sheets, tags & roles, season wheel, desert/qanat procedures, caravan/oasis/council scores, oath & bloodŌłÆsilver economy, generators, integration, and an example.

\textbf{End of v0.1}


\section*{FHARA ŌłÆŌłÆŌłÆ Caravan Law & LampŌłÆLit Bargains}

\textbf{Where you'll meet them:} Oasis belts, tentŌłÆcourts, caravanserai streets, night markets, relief trains.

\textbf{Strings (pick 2ŌłÆŌłÆ3 to start):} waterŌłÆshare tablets $ŌĆó$ caravan court writs $ŌĆó$ escort oaths $ŌĆó$ saltŌłÆroad passes.

\textbf{Etiquette Hooks}
\begin{itemize}
Ōä╣tem \textbf{Water First:} Offer water before business ŌåÆ Position +1 on Petition/Broker.
Ōä╣tem \textbf{Lamp for Contracts:} Light a small lamp for signatures ŌåÆ +1 Effect on a contractŌłÆbinding roll at dusk courts.
Ōä╣tem \textbf{Verse of Courtesy:} Open with a short couplet of thanks ŌåÆ convert Audience: Cool ŌåÆ Warm with officials.
\end{itemize}

\textbf{Faux Pas (SB):} Parched Pride ŌłÆŌłÆŌłÆ skipping water service flips Audience against you.

\textbf{Venue Tags}
\begin{itemize}
Ōä╣tem \textbf{Caravan Court:} neutral tent tribunal; LampŌłÆOaths bind as durable Strings; breaking ticks Exposure +1.
Ōä╣tem \textbf{Moon Market:} night bazaar; Market starts Position +1; on 1, rumor spreads (Lamplighter Gossip).
\end{itemize}

\textbf{Mounts & Gear}
\begin{itemize}
Ōä╣tem \textbf{Desert Camel:} WaterŌłÆwise, HeatŌłÆhardy (prevent one thirst Supplies ŌłÆŌłÆ1 per leg).
Ōä╣tem \textbf{Covered Wagon:} grants shade; first Heat tick/day ignored while traveling slowly.
\end{itemize}

\textbf{Common Play Roles}\\
Envoy, Quartermaster, WaterŌłÆFinder, LampŌłÆNotary, Escort Captain, PoetŌłÆMediator.

\textbf{Factions}\\
Caravan Courts $ŌĆó$ WaterŌłÆShare Clerks $ŌĆó$ LampŌłÆGuilds $ŌĆó$ Escort Companies $ŌĆó$ Shrine Hosts.

\textbf{Adventure Seeds}
\begin{itemize}
Ōä╣tem A forged waterŌłÆshare tablet triggers Oasis Water Rights Dispute [6].
Ōä╣tem An escort oath is broken; recover cargo or stand surety at midnight court.
\end{itemize}

\section*{SIDHI ŌłÆŌłÆŌłÆ Pilots, Traders, and CoastŌłÆRoad Courts}

\textbf{Where you'll meet them:} Coastal entrep├┤ts, river mouths, lighthouse towns, caravan junctions, innerŌłÆquarter courts.

\textbf{Strings:} pilot stones $ŌĆó$ ferry contracts $ŌĆó$ lighthouse charters $ŌĆó$ guild guestŌłÆrights.

\textbf{Etiquette Hooks}
\begin{itemize}
Ōä╣tem \textbf{Bread & Salt:} Share a simple plate before negotiation ŌåÆ Position +1 in Parley at homes or guild halls.
Ōä╣tem \textbf{Name the Line:} Give family/mentor line when introduced ŌåÆ Audience: Cool ŌåÆ Warm with civic officials.
Ōä╣tem \textbf{GuestŌłÆRight Token:} Present a stamped guestŌłÆright from a guild ŌåÆ DV ŌłÆŌłÆ1 on first Petition in that venue.
\end{itemize}

\textbf{Faux Pas (SB):} Cut Line ŌłÆŌłÆŌłÆ ignoring lineage or guestŌłÆright protocol imposes Position ŌłÆŌłÆ1 for the scene.

\textbf{Venue Tags}
\begin{itemize}
Ōä╣tem \textbf{Pilots' Quay:} navigation disputes settled by local charts; Clue +1 on routes with a token.
Ōä╣tem \textbf{Lighthouse Court:} tolls and wreck law; Board & Brace DV ŌłÆŌłÆ1 when acting as sworn rescue.
\end{itemize}

\textbf{Mounts & Gear}
\begin{itemize}
Ōä╣tem \textbf{River Barge Teams:} Riverwise; Row Upstream DV ŌłÆŌłÆ1 with trained crews.
Ōä╣tem \textbf{Pilot Stones Map:} once/leg Navigate Position +1 through shoals.
\end{itemize}

\textbf{Common Play Roles}\\
Factor, Pilot, WreckŌłÆWarden, Guild Mediator, Street Broker, Quay Advocate.

\textbf{Factions}\\
Pilots' Union $ŌĆó$ Lighthouse Wardens $ŌĆó$ Quay Guilds $ŌĆó$ Ferry Syndicates $ŌĆó$ Shrine of Safe Crossing.

\textbf{Adventure Seeds}
\begin{itemize}
Ōä╣tem A lighthouse charter is forged; expose the plot before a storm convoy arrives.
Ōä╣tem Pilots' stones are moved to benefit wreckers; restore the line under cover of fog.
\end{itemize}

\section*{ASHAANI ŌłÆŌłÆŌłÆ Processions, Weighhouses, and Cistern Law}

\textbf{Where you'll meet them:} Provincial capitals, desert watchposts, processional avenues, cistern towns, shrine courts.

\textbf{Strings:} weighhouse seals $ŌĆó$ festival licenses $ŌĆó$ relief priorities $ŌĆó$ cistern ledgers.

\textbf{Etiquette Hooks}
\begin{itemize}
Ōä╣tem \textbf{Titles Before Names:} Use offices first ŌåÆ Audience: Cool ŌåÆ Warm with administrators.
Ōä╣tem \textbf{Procession Deference:} Step aside, lower voices during rites ŌåÆ Position +1 on the next roll with witnesses.
Ōä╣tem \textbf{Seal & Receipt:} Present stamped receipt at weighhouse ŌåÆ DV ŌłÆŌłÆ1 on audits/inspections.
\end{itemize}

\textbf{Faux Pas (SB):} Defaced Seal ŌłÆŌłÆŌłÆ mishandling a seal lets the GM start Inspection [4].

\textbf{Venue Tags}
\begin{itemize}
Ōä╣tem \textbf{Weighhouse:} tariff hall; on 1, Inspection [4] advances; exact tallies can convert Exposure ŌłÆŌłÆ1.
Ōä╣tem \textbf{Processional Way:} festivals; Crowd tag can be spent for Position +1 in public persuasion.
\end{itemize}

\textbf{Mounts & Gear}
\begin{itemize}
Ōä╣tem \textbf{Relief Cart Trains:} Burdened, ShadeŌłÆCanopy; Heat ŌłÆŌłÆ1 on rest if canopies are intact.
Ōä╣tem \textbf{Ledger Clips:} prove payment; once/stop convert Inspection ŌłÆŌłÆ1.
\end{itemize}

\textbf{Common Play Roles}\\
Auditor, Quartermaster, Procession Marshal, Shrine Scribe, Road Captain.

\textbf{Factions}\\
Satrap Offices $ŌĆó$ Relief Granaries $ŌĆó$ Processional Guilds $ŌĆó$ Shrine Courtiers $ŌĆó$ Desert Watch.

\textbf{Adventure Seeds}
\begin{itemize}
Ōä╣tem Relief grain diverted by a weighhouse clerk; audit under pressure amid a festival.
Ōä╣tem A procession mask is stolen; retrieve it before rites fail and Crisis ticks up.
\end{itemize}

\section*{PERESHI ŌłÆŌłÆŌłÆ CharterŌłÆMakers, Gardeners, and HighŌłÆRoad Hosts}

\textbf{Where you'll meet them:} Inland caravan routes, terrace towns, archive courts, roofŌłÆgardens, mountain passes.

\textbf{Strings:} landŌłÆcharter tablets $ŌĆó$ archive citations $ŌĆó$ highŌłÆroad guestŌłÆrings $ŌĆó$ garden waterŌłÆshares.

\textbf{Etiquette Hooks}
\begin{itemize}
Ōä╣tem \textbf{Fruit & Tea:} Offer fruit and a warm cup on arrival ŌåÆ Position +1 in household negotiations.
Ōä╣tem \textbf{Shoes at Threshold:} Observe house thresholds; remove shoes indoors ŌåÆ negate a Position ŌłÆŌłÆ1 penalty from prior brusqueness.
Ōä╣tem \textbf{Couplet of Courtesy:} A brief couplet acknowledging hosts' craft ŌåÆ Audience: Cool ŌåÆ Warm with artisans/scribes.
\end{itemize}

\textbf{Faux Pas (SB):} Dust on the Mat ŌłÆŌłÆŌłÆ barging in tracks Exposure +1 among neighbors.

\textbf{Venue Tags}
\begin{itemize}
Ōä╣tem \textbf{Archive Court:} precedentŌłÆdriven; present citations to treat a social DV ŌłÆŌłÆ1 once/scene.
Ōä╣tem \textbf{RoofŌłÆGarden Pact:} settlements on rooftops; a hosted pact counts as a String across partnered houses.
\end{itemize}

\textbf{Mounts & Gear}
\begin{itemize}
Ōä╣tem \textbf{Pack Mules:} SureŌłÆfooted; Cross (switchbacks) DV ŌłÆŌłÆ1.
Ōä╣tem \textbf{HighŌłÆRoad Litter:} sheltered; grants Heat ŌłÆŌłÆ1 on long climbs at the cost of Distance ŌłÆŌłÆ1 progress/day.
\end{itemize}

\textbf{Common Play Roles}\\
Land Clerk, HighŌłÆRoad Host, Charter Advocate, Garden Steward, Archive Scribe, Pass Warden.

\textbf{Factions}\\
Archive Leagues $ŌĆó$ Gardeners' Courts $ŌĆó$ HighŌłÆRoad Hospices $ŌĆó$ CharterŌłÆMakers $ŌĆó$ Terrace Wardens.

\textbf{Adventure Seeds}
\begin{itemize}
Ōä╣tem Two terrace towns feud over a spring; arbitrate via archive citations and garden pacts.
Ōä╣tem A forged charter threatens to displace a hospice; uncover the scribe ring.
\end{itemize}

\section*{KUVANI ŌłÆŌłÆŌłÆ Steppe Hosts, Remount Chains, and SkyŌłÆRoad Songs}

\textbf{Where you'll meet them:} Meadow edges, steppe roads, remount posts, border fairs, winter camps.

\textbf{Strings:} remount chain marks $ŌĆó$ steppe guestŌłÆropes $ŌĆó$ grazing compacts $ŌĆó$ skyŌłÆroad songs (route lore).

\textbf{Etiquette Hooks}
\begin{itemize}
Ōä╣tem \textbf{Guest Tent Rope:} Leave a rope coiled outside to signal neutrality ŌåÆ Position +1 on first Parley in camp.
Ōä╣tem \textbf{Bowl Before Blade:} Accept a simple bowl before speaking of arms ŌåÆ Audience: Cool ŌåÆ Warm with elders.
Ōä╣tem \textbf{Honor the Horse:} Praise mounts; tend tack before sleep ŌåÆ DV ŌłÆŌłÆ1 on next Chase related to horse teams.
\end{itemize}

\textbf{Faux Pas (SB):} Spurned Mount ŌłÆŌłÆŌłÆ neglecting a horse lets the GM impose Position ŌłÆŌłÆ1 among riders.

\textbf{Venue Tags}
\begin{itemize}
Ōä╣tem \textbf{Remount Post:} swap fresh mounts; once/leg treat Chase (horse) DV ŌłÆŌłÆ1.
Ōä╣tem \textbf{SkyŌłÆRoad Cairns:} waystones; spend a song token to convert Navigate Position +1 across open grass.
\end{itemize}

\textbf{Mounts & Gear}
\begin{itemize}
Ōä╣tem \textbf{Remount String:} rotating horses; ignore first Weariness +1 from a Chase each day.
Ōä╣tem \textbf{Steppe Bow Kit:} in open terrain gain +1 Effect on first ranged volley of a clash.
\end{itemize}

\textbf{Common Play Roles}\\
HostŌłÆCaptain, Scout, LoreŌłÆKeeper, Herd Steward, Banner Rider, Border Mediator.

\textbf{Factions}\\
Remount Leagues $ŌĆó$ RiverŌłÆEdge Hosts $ŌĆó$ Banner Circles $ŌĆó$ Steppe Courts $ŌĆó$ WayŌłÆShrine Keepers.

\textbf{Adventure Seeds}
\begin{itemize}
Ōä╣tem Grazing compacts broken by a border satrap; negotiate new strings before herds clash.
Ōä╣tem A remount post is burned; rebuild the chain under threat from raiders.
\end{itemize}

\section{CrossŌłÆPeoples Integration (quick rules)}
\begin{itemize}
Ōä╣tem \textbf{Etiquette Equivalence:} Any culture's etiquette hook can substitute in mixed venues if a local sponsor vouches; treat as Position +0 (not +1) unless the PCs hold a relevant String.
Ōä╣tem \textbf{Strings as Passports:} A String earned in one sphere (e.g., waterŌłÆshare tablet) lowers DV or raises Position in another when fictionally relevant (e.g., purchasing feed at a Sidhi quay).
Ōä╣tem \textbf{Markets:} Map appetites across peoplesŌłÆŌłÆŌłÆFhara (Relief/Ritual), Sidhi (Harbor/Luxury), Ashaani (Official/Relief), Pereshi (Artisan/Scholarly), Kuvani (Travel/Beast). Apply +1 Position when selling into a matching appetite with a native broker.
Ōä╣tem \textbf{Followers:} When recruiting followers from these peoples, give one culture tag that can be spent once/leg for Position +1 (e.g., WaterŌłÆFinder, Pilot, Procession Marshal, Archive Clerk, Remount Handler).
\end{itemize}

\section{GM Portrayal Notes (keep it plural)}
\begin{itemize}
Ōä╣tem Represent multiple livelihoods within each people (clerks and camelŌłÆdrivers; pilots and street brokers; gardeners and charterŌłÆmakers; riders and loreŌłÆkeepers).
Ōä╣tem Use oaths, courts, and hospitality to show social fabric; let etiquette shift mechanics, not caricature.
Ōä╣tem Ask players what their characters learned from hosts; bring that forward as advantage tags and Strings.
Ōä╣tem When in doubt, lean on law and reciprocity over spectacle.
\end{itemize}

\section*{Print Aid}

Arrange each culture as a halfŌłÆpage card. Add icons: cup (hospitality), lamp (oaths), scale (law), mount (travel), knot (Strings), wave/grass/brick for venue type.

\section*{Changelog}

\textbf{v0.1} ŌłÆŌłÆŌłÆ Five quick cards with etiquette hooks, lawful Strings, venue tags, mounts/gear, roles, factions, seeds, crossŌłÆintegration, and portrayal notes.

\textbf{End of v0.1}

% ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ ADVENTURE 1 ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ
\section*{The Thirst Ledger (Tier II)}
\textit{Oasis arbitration, qanat sabotage, and lampŌłÆcourt verdicts.}

\paragraph{Overview} In the Fharan oasis of \textbf{Bir alŌłÆQamar}, a forged waterŌłÆshare tablet ignites a crisis. The crew must audit ledgers in the \emph{Qanat Quarter}, unmask saboteurs in the wadis, and argue terms before the dusk \emph{Lamp Court}.

\paragraph{Peoples at the Table} 
\textbf{Fhara} caravan court clerks; \textbf{Sidhi} quay factors with stored cistern grain; \textbf{Ashaani} weighhouse auditors in transit; \textbf{Pereshi} archiveŌłÆcitation scribe; \textbf{Kuvani} remount outriders guarding relief carts.

\subsection*{Fronts & Timers}
\begin{itemize}
  Ōä╣tem \textbf{Oasis Water Rights Dispute [6]} ŌłÆŌłÆŌłÆ riots if unresolved.
  Ōä╣tem \textbf{Qanat Warden Alert [4ŌłÆŌłÆ6]} ŌłÆŌłÆŌłÆ patrols tighten; Exposure on miss.
  Ōä╣tem \textbf{LampŌłÆCourt Patience [4]} ŌłÆŌłÆŌłÆ judge ends hearings if filled.
  Ōä╣tem \textbf{Khamsin Front [4]} ŌłÆŌłÆŌłÆ dust gales raise Heat/Condition.
\end{itemize}

\subsection*{Key Sites (Tags)}
\textbf{Moon Market} (Market +1 Position; SB: Lamplighter Gossip), 
\textbf{Qanat Quarter} (Infiltrate +1 Position; on 1: Exposure +1), 
\textbf{Lamp Court} (LampŌłÆOaths bind), 
\textbf{Wadi Country} (Stealth +1 Position; Flash Flood risk).

\subsection*{Beats & Procedures}
\textbf{B1: Market Murmurs} (Presence+Sway, DV 3). Etiquette: \emph{Water First} $ŌåÆ$ Position +1. On 1: SB \emph{Lamplighter Gossip} flips Audience. \\
\textbf{B2: Qanat Audit} (Wits+Diplomacy/Audit, DV 3ŌłÆŌłÆ4). On strong hit, reveal ledger splice; on 1, start \emph{Inspection [4]}. \\
\textbf{B3: Wadi Pursuit} (Chase: Foot/Beast DV 2ŌłÆŌłÆ4 by terrain). \emph{Kuvani Remount Post} lowers DV by 1 once/leg. \\
\textbf{B4: LampŌłÆCourt Plea} (Presence+Sway, DV 3). \emph{LampŌłÆlit Contract} grants +1 Effect; stake a \emph{LampŌłÆOath String} with season timer [4ŌłÆŌłÆ6].

\subsection*{Hazards & Sirocco SB}
\emph{Thirst Bites} (Heat +1), \emph{Mirage Cut} (Distance stalls), \emph{Qanat Quarrel} (Inspection [4]), \emph{Khamsin} (Supplies ŌłÆŌłÆ1; Condition +1).

\subsection*{Outcomes & Rewards}
If the crew proves the forgery: reduce \emph{Oasis Water Rights Dispute} to 0; gain \textbf{Strings}: waterŌłÆshare tablet, tentŌłÆcourt seal; Standing with caravan courts or wardens. Failure ticks Crisis and spawns a riot scene.

\subsection*{Integration Hooks}
Political Intrigue (\emph{LampŌłÆOath Engine}); Caravans (Escort relief carts under Heat); Wilderness (Flash Flood crossing); Sea/Sidhi (quay granaries).

% ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ ADVENTURE 2 ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ
\section*{Night of the Moon Kilns (Tier IIŌłÆŌłÆIII)}
\textit{A glass monopoly war in Sar Qalyat's wind towers.}

\paragraph{Overview} In \textbf{Sar Qalyat}, the Galaninan Mirror League runs a secret moonŌłÆkiln with outlawed sand. Lamplighters strike; wardŌłÆlines whisper. The crew can heist the kiln, flip contracts in lampŌłÆcourt, or expose saboteurs.

\paragraph{Peoples at the Table}
\textbf{Fhara} envoy brokering night freight; \textbf{Sidhi} factor with harbor buyers; \textbf{Ashaani} courier bearing festival seals; \textbf{Pereshi} glassŌłÆmeasurer; \textbf{Kuvani} dustŌłÆsailer crew hired as decoy.

\subsection*{Fronts & Timers}
\begin{itemize}
  Ōä╣tem \textbf{Mirror War [6ŌłÆŌłÆ8]} ŌłÆŌłÆŌłÆ assassins, sabotage, price shocks.
  Ōä╣tem \textbf{Lamplighter Strike [4]} ŌłÆŌłÆŌłÆ escorts needed to move safely by night.
  Ōä╣tem \textbf{WardŌłÆLine Murmurs [6]} ŌłÆŌłÆŌłÆ wardŌłÆstorm risk rises near ruins quarter.
  Ōä╣tem \textbf{Inspection Dragnet [6]} ŌłÆŌłÆŌłÆ weighhouse sweeps for contraband sand.
\end{itemize}

\subsection*{Key Sites (Tags)}
\textbf{Wind Towers} (Heat advances slower; Navigate rooftop +1 Position), 
\textbf{Moon Market} (Market +1 Position), 
\textbf{Lamp Court} (+1 Effect with lampŌłÆgift),
\textbf{Ruin Fields Edge} (WardŌłÆStorm risk).

\subsection*{Beats & Procedures}
\textbf{B1: Rooftop Recon} (Wits+Survey, DV 3). Pilot lines, watch rotations; on 1: SB \emph{Witness at Roof} starts \emph{Inspection [4]}.\\
\textbf{B2: Contract Turn} (Presence+Sway, DV 3). \emph{LampŌłÆlit Contract} grants +1 Effect; a Pereshi citation reduces DV ŌłÆŌłÆ1 once/scene.\\
\textbf{B3: MoonŌłÆKiln Heist} (MultiŌłÆaction: Infiltrate DV 3, Disable DV 4, Escape DV 3). DustŌłÆsailer decoy $ŌåÆ$ \emph{Chase (dust)} DV ladder 2ŌłÆŌłÆ5.\\
\textbf{B4: WardŌłÆStorm Crossing} (Psionics opt.). Without wards, scene starts \emph{Desperate}; strong hit clears Heat/Morale ŌłÆŌłÆ1.

\subsection*{Hazards & SB}
\emph{Lamplighter Gossip} (Audience flips), 
\emph{DustŌłÆSail Ambush} (start Pursuit [6]), 
\emph{Ward Surge} (Scene Desperate), 
\emph{Inspection Sweep} (Dragnet +1).

\subsection*{Outcomes & Rewards}
Strings: lamplighter contract, mirrorŌłÆtoll beads, ward indemnity. Choose: shutter the illegal kiln (Standing with lamplighters) or seize its contracts (coin + Exposure). Sidhi factor can open sea lanes if favored.

\subsection*{Integration Hooks}
Caravans (night freight), Psionics (WardŌłÆStorm), Sea (Sidhi buyers), Political (court injunctions), Dragons (mirrored halls in ruin fields).

% ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ ADVENTURE 3 ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ
\section*{Breath of Relief (Tier III)}
\textit{Escort a relief train through khamsin and raids to the Qattara Lines.}

\paragraph{Overview} Famine grips the Ashaani provinces. A relief caravan must cross salt flats, survive dust gales, and pass a crooked weighhouse. Raiders and wardŌłÆstorms threaten every leg.

\paragraph{Peoples at the Table}
\textbf{Fhara} escort captains and court notaries; \textbf{Sidhi} pilots coordinating river offload; \textbf{Ashaani} satrap officials and procession marshals; \textbf{Pereshi} auditors tracking grain; \textbf{Kuvani} outriders screening the regs.

\subsection*{Fronts & Timers}
\begin{itemize}
  Ōä╣tem \textbf{Relief Train in Peril [6]} ŌłÆŌłÆŌłÆ if filled, famine riots tick Crisis.
  Ōä╣tem \textbf{Sirocco Raiders [6]} ŌłÆŌłÆŌłÆ dustŌłÆsail ambushes along flats.
  Ōä╣tem \textbf{WardŌłÆStorm Front [4]} ŌłÆŌłÆŌłÆ psionic weather spikes.
  Ōä╣tem \textbf{Weighhouse Corruption [4]} ŌłÆŌłÆŌłÆ false tallies, bribe attempts.
\end{itemize}

\subsection*{Leg Structure (use Intent Dials)}
\textbf{Leg A: Salt Flats (Speed)} ŌłÆŌłÆŌłÆ strong hits add Distance +1; on 1: Supplies ŌłÆŌłÆ1, Condition +1. \\
\textbf{Leg B: Wadis (Stealth)} ŌłÆŌłÆŌłÆ avoid Pursuit; Flash Flood check (Cross DV 4ŌłÆŌłÆ5). \\
\textbf{Leg C: Ruin Fields (Endure)} ŌłÆŌłÆŌłÆ WardŌłÆStorm scene; wards negate Desperate. \\
\textbf{Leg D: Weighhouse (Show)} ŌłÆŌłÆŌłÆ Procession etiquette $ŌåÆ$ Position +1; Audit DV 3.

\subsection*{Key Scenes}
\textbf{S1: Khamsin Wall} (Body+Tactics, DV 3). Quartermaster converts Supplies $ŌåÆ$ Heat ŌłÆŌłÆ1 once/day. \\
\textbf{S2: DustŌłÆSail Pursuit} (Chase DV: 2 flats ŌĆó 3 gusty ŌĆó 4 crosswind ŌĆó 5 khamsin). Kuvani remount chain can substitute once (treat as longship mapping if using Linn tools). \\
\textbf{S3: Weighhouse Trial} (Presence+Sway/Audit, DV 3). TitlesŌłÆbeforeŌłÆnames converts Audience; Pereshi ledger clips reduce \emph{Inspection} by 1.

\subsection*{Sirocco SB Menu}
\emph{Thirst Bites}, \emph{Mirage Cut}, \emph{DustŌłÆSail Ambush}, \emph{Lamplighter Gossip} (at night market), \emph{Ward Surge}.

\subsection*{Outcomes & Rewards}
Deliver relief: Standing with satrap office and caravan courts; Strings: festival license, weighhouse seal, escort oath. Failure spawns \emph{Relief Riot [4]} and reduces regional Mandate.

\subsection*{Integration Hooks}
Political (mandate/relief), Caravans (orders, Overextend), Wilderness (Heat/Camp), Sea (Sidhi offload at port).

% ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ ADVENTURE 4 ŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆŌłÆ
\section*{Song Beneath the Dunes (Tier IIIŌłÆŌłÆIV)}
\textit{WardŌłÆlines awaken under a mirrored ruin; bargains with dustŌłÆdjanni.}

\paragraph{Overview} RuinŌłÆkeepers hire the crew to stabilize a wardŌłÆline below a duneŌłÆshrouded ziggurat. A rival dig seeks forbidden sand. The crew must cross by star paths, parley with a dustŌłÆdjanni, and decide what power the wards will serve.

\paragraph{Peoples at the Table}
\textbf{Fhara} guides and waterŌłÆfinders; \textbf{Sidhi} artifact brokers awaiting port sale; \textbf{Ashaani} shrine courtiers with ritual indemnities; \textbf{Pereshi} scholars mapping star shafts; \textbf{Kuvani} scouts screening raiders.

\subsection*{Fronts & Timers}
\begin{itemize}
  Ōä╣tem \textbf{Awakening WardŌłÆLines [8]} ŌłÆŌłÆŌłÆ scenes trend Desperate; psionic echoes.
  Ōä╣tem \textbf{Rival Digs [6]} ŌłÆŌłÆŌłÆ sabotage, poaching workers.
  Ōä╣tem \textbf{Djanni Bargain [4]} ŌłÆŌłÆŌłÆ unmet terms bring Heat +1/leg.
  Ōä╣tem \textbf{Heat Track [6]} ŌłÆŌłÆŌłÆ if full, Harm +1 then reset to 3.
\end{itemize}

\subsection*{Key Sites (Tags)}
\textbf{Erg Dunes} (Navigate DV +1 without stars), 
\textbf{Star Shaft} (LoreŌłÆKeeper lowers DV ŌłÆŌłÆ1 at night), 
\textbf{Mirrored Hall} (WardŌłÆStorm risk), 
\textbf{Shrine of the North Star} (Navigate DV ŌłÆŌłÆ1 once/leg).

\subsection*{Beats & Procedures}
\textbf{B1: Night Approach} (Star Paths, DV ŌłÆ1 with astrolabe). Strong hit grants a Clue to ward nodes. \\
\textbf{B2: Djanni Parley} (Presence+Rites, DV 3). Offer verse+water to gain \emph{Passage String}; on miss, \emph{Djanni Bargain} advances. \\
\textbf{B3: Stabilize the Line} (Wits+Craft/Psionics, DV 4). On strong hit, reduce \emph{Awakening WardŌłÆLines ŌłÆ2}. On 1, scene becomes Desperate; Fatigue risk. \\
\textbf{B4: Rival Interference} (Chase or Skirmish). Use Regs for Ambush +1 Position; capture their ledger for exposure in lampŌłÆcourt later.

\subsection*{Hazards & SB}
\emph{Khamsin}, \emph{Ward Surge}, \emph{Mirage Cut}, \emph{Qanat Quarrel} (supply spat at camp).

\subsection*{Outcomes & Rewards}
Choose one: bind wards to shrine (Standing with ruinŌłÆkeepers; sacred cargo boon), redirect to city grid (Standing with Mirror League; coin), or slacken the line to map a dragon route (clue to hoard; Exposure +1). Strings: wardŌłÆring, oathŌłÆlamp, survey map.

\subsection*{Integration Hooks}
Psionics (Aegis, Mental Navigation), Political (who controls the ward grid), Dragons (ancient routes), Caravans (sacred cargo transport), Sea (Sidhi brokers).

\end{document}