\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{array}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{tikz}
\usepackage{framed}
\usepackage{xcolor}

\geometry{margin=1in}
\setstretch{1.15}
\pagestyle{fancy}
\fancyhf{}
\rhead{Fate's Edge Module}
\lhead{Shadows \& Steel v0.2}
\rfoot{\thepage}

\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}

\newcommand{\moduletitle}{Shadows \& Steel — Underworld and Criminal Networks (Fate's Edge v0.2)}
\newcommand{\subtitle}{A criminal‑campaign module for Fate's Edge: crime families, smugglers, black markets, fences, fixers, and the law that hunts them. Built to sit beside Violets \& Stone (districts), Political Intrigue (faction play), Caravans (smuggling legs), Allies \& Adversaries (crew/org management), Assets \& Patrons, and the core SRD. Uses Position/DV, SB, clocks, Strings, Favor/Leverage/Exposure—no new math.}

\begin{document}

\begin{center}
{\LARGE \textbf{Shadows \& Steel — Underworld and Criminal Networks}} \\
{\large (Fate's Edge v0.2)}
\end{center}

\vspace{0.2in}

A criminal‑campaign module for Fate's Edge: crime families, smugglers, black markets, fences, fixers, and the law that hunts them. Built to sit beside Violets \& Stone (districts), Political Intrigue (faction play), Caravans (smuggling legs), Allies \& Adversaries (crew/org management), Assets \& Patrons, and the core SRD. Uses Position/DV, SB, clocks, Strings, Favor/Leverage/Exposure—no new math.

\section*{What's New in v0.2 (from playtest)}
\begin{itemize}
\item More playbooks: added Fixers and Ward‑Breakers.
\item Heat threshold table: formal Crew/City Heat triggers; Setup SB → Heat/Alarm rule.
\item Overhead tools: Clock Budget (3+1), Heat‑to‑Complication conversion, one‑page Heist Cheat Sheet.
\item Supernatural tie‑ins: Arcane Tier black market, ward crimes, psionic hooks.
\item Long arcs: three campaign frames with milestones.
\item Political mapping: War ↔ Faction clocks, formal Blood‑Silver + Witness settlement.
\end{itemize}

\section*{Quickstart (2 minutes)}
\begin{enumerate}
\item Make a Crew Sheet (§1) and pick a Playbook (Smugglers, Thieves, Enforcers, Grifters, Fixers, Ward‑Breakers). Choose 3 Crew Tags and 2 Strings.
\item Sketch a Turf Map (§2) with 4–6 nodes (districts/venues/routes). Mark 1 Racket you control.
\item Start tracks: Crew Heat [6], City Heat [6], Notoriety [4] (opt.), Law Mandate/Crisis 2/2.
\item Choose an opening Score Type (§6): Heist, Smuggling Run, Protection Play, or Confidence Game.
\item On any 1, GM spends from Shadows \& Steel SB (§9.4). Setup Rule: each SB gained during Setup may either Alarm +1 or Crew Heat +1 (GM picks by fiction).
\end{enumerate}

\section{Crew Sheet (Template)}
\begin{framed}
\noindent \textbf{[CREW NAME]} \\
Type: Smugglers / Thieves / Enforcers / Grifters / Fixers / Ward‑Breakers \\
Tags (choose 3): Riverwise • Locksmiths • Shadow Accounts • Bribe Web • Knives Out • Silver Tongues • Wrecker's Fire • Clean Faces • Ghost Safehouses • Ward Keys • Court Whisper \\
Strings (choose 2): dockmaster favor • tollhouse key • guild fence • customs blind • patrol captain's debt • theater front • lamp‑notary on retainer • ruin‑keeper indulgence \\
Tracks:
\begin{itemize}
\item Crew Heat [6] (recent attention)
\item Notoriety [4] (opt., famous style)
\item Rep [6] (standing underground)
\item Bank (coin abstraction)
\end{itemize}
Allies/Rivals: fixers, fences, patrol captains, rival crews \\
Safehouses/Fronts: names + tags
\end{framed}

\subsection*{Crew Benefits by Type}
\begin{itemize}
\item \textbf{Smugglers:} start Riverwise, Customs Blind; once/leg, Smuggle DV –1.
\item \textbf{Thieves:} start Locksmiths, Clean Faces; once/score, Infiltrate Position +1.
\item \textbf{Enforcers:} start Knives Out, Wrecker's Fire; once/score, Intimidate/Break Effect +1.
\item \textbf{Grifters:} start Silver Tongues, Shadow Accounts; once/score, convert Audience: Cool → Warm.
\item \textbf{Fixers (NEW):} start Bribe Web, Court Whisper; once/score, treat Petition/Broker DV –1 in courts/markets.
\item \textbf{Ward‑Breakers (NEW):} start Ward Keys, Ghost Safehouses; once/score, ignore first Desperate from wards/psionic weather.
\end{itemize}

\section{Turf Map \& Territory Control}
Turf Map: Draw 4–6 nodes: Districts (e.g., Dye Yards, Pilots' Quays), Venues (Moon Market, Gambling Row), Pipelines (river cut, sewer run, postern stairs). Connect with arrows that show movement/contraband flow.

\subsection*{Rackets:} gambling den • dice house • ropewalk skims • dock tithe • counterfeit inks • lanternhouses • fighting pits • lockshop chain • locker barge • eel‑channel • ward‑glass salvage (NEW) • oath‑lamp forgery (NEW).

\subsection*{Taking Turf:} create Claim [6] on a node. During scores aimed at this node, strong outcomes tick Claim. When it fills, add the racket and a String (front or bribe web). Mark Crew Heat +1.

\subsection*{Holding Turf:} each session, roll Hold (Wits+Tactics) vs DV 2–4 by pressure. On hit, gain Bank +1 or Audience: Warm in that node. On 1s, advance a Trouble [4] for that turf (rivals, inspectors, unions).

\subsection*{Territory War:} When rivals contest, start War [8]. While War>0, Bandwidth –1 between sessions. Settlement Template: pay Blood‑Silver (2 Bank each) + name a Witness Faction; all parties stake a Lamp‑Oath [4]. Breaking it: Exposure +1, Repute –1 with the witness faction, War resumes at 2/8.

\section{Heat, Law, and Crackdowns}
\subsection*{Crew Heat [6] thresholds}
\begin{itemize}
\item 0–2: background hum.
\item 3: start Patrol Sweep [4].
\item 4–5: random stops; Notoriety +1 when you make a flashy move.
\item 6: trigger Sting [6] and Exposure +1.
\end{itemize}

\subsection*{City Heat [6] thresholds}
\begin{itemize}
\item 0–2: routine.
\item 3: curfew murmurs, stop‑and‑search; start Patrol Sweep [4] in hot districts.
\item 6: Crackdown [4] (DV +1 on criminal actions, law starts with Audience: Hostile).
\end{itemize}

\subsection*{Cooling Actions}
\begin{itemize}
\item \textbf{Lay Low:} lose 1 Bandwidth; clear Crew Heat –2; gain Notoriety +1 (opt.).
\item \textbf{Grease Palms:} spend Bank/Favor; clear Crew Heat –1 and one Inspection.
\item \textbf{Hard Pivot:} burn a front (lose one racket) to clear Crew Heat –3 and City Heat –1.
\end{itemize}

Law Tools: Riverine Patrol, City Watch, Tax Assessor, Bridge Guards—each tracks a Sting [6].

\section{Black Markets \& Rare Goods}
Use Market from Caravans.

\subsection*{Tiers \& DV:} Street DV 2–3 • Guild DV 3–4 • Court DV 4–5 • Arcane (NEW) DV 5 (wards, psionic keys, oath‑lamps). \\
Tags: Contraband, Hot, Forged, Sacred, Perishable, Traceable, Warded (NEW).

\subsection*{Partials:} price spike (Bank –1) • Crew Heat +1 • quality flaw (–1 Effect next use) • seller's Temp String • Inspection [4] starts. \\
\textbf{Miss:} sting springs (Sting [6] +2), counterfeit, or owed favor to a rival fence. \\
\textbf{Laundering:} Broker/Account DV 3–4. On hit, convert Hot → Clean; on 1, City Heat +1.

\section{Informant Webs}
Map a Web: 3–5 Informants (dock porter, lamplighter, clerk, patrol runner, skald…). Each has Access, Reliability [4], Price, Flags.

Use: pay Price for Clue +1, small String, or DV –1 in their sphere. On 1s, Reliability –1. At 0, they flip or vanish; start Burn [4] to recover/replace.

\subsection*{Moves (NEW):}
\begin{itemize}
\item \textbf{Cross‑Check:} spend 2 informants on same topic → take +1 Effect and only the lowest Reliability –1 applies on a 1.
\item \textbf{Mask a Source:} spend Bank 1 to prevent Reliability –1 once.
\end{itemize}

\section{Score Types \& Heist Planner}
\subsection*{Approaches:} Deceit • Stealth • Violence • Bribe • Con.

# Heist System for Shadows and Steel

\subsection{Heist Operations}

\subsubsection{Heist Framework}

Heists in the shadows of Fate's Edge require careful planning, precise execution, and the ability to adapt when things go wrong. The following framework provides structure while maintaining the flexibility essential to underground operations.

\subsubsection{Heist Phases}

\begin{description}
    \item[Planning] Intelligence gathering, approach vector selection, and resource allocation
    \item[Setup] Positioning, final preparations, and initial infiltration
    \item[Execution] Core operation with primary objectives
    \item[Extraction] Escape and aftermath management
\end{description}

\subsubsection{Heist Planner}

Before any heist, complete this essential planning document:

\begin{framed}
\textbf{Heist Planner Template}

\textbf{Objective:} \rule{4cm}{0.1pt}

\textbf{Venue:} \rule{4cm}{0.1pt}

\textbf{Entry/Exit Strings:} 
\begin{itemize}
    \item \rule{5cm}{0.1pt}
    \item \rule{5cm}{0.1pt}
\end{itemize}

\textbf{Setup Rolls (0-2):} Each strong hit grants Position +1 or DV -1 later; each 1 → GM +1 SB or Alarm +1/Crew Heat +1

\textbf{Threat Clocks:} 
\begin{itemize}
    \item Alarm [4-6] \rule{3cm}{0.1pt}
    \item Patrol Routes [4] \rule{3cm}{0.1pt}
    \item Inside Man [4] \rule{3cm}{0.1pt}
\end{itemize}

\textbf{Flashback Opportunities:} \rule{6cm}{0.1pt}
\end{framed}

\subsubsection{Kit Tokens}

Each crew member begins a heist with 2 Kit Tokens representing plausible gear or contacts. Additional tokens beyond 2 add Crew Heat +1.

\begin{description}
    \item[Lock Picks] Precision tools for mechanical security
    \item[Dark Lantern] Concealable illumination
    \item[Silk Rope] Lightweight climbing equipment
    \item[False Papers] Identity documents for social infiltration
    \item[Safe House Key] Emergency refuge access
    \item[Informant Favor] One-time assistance from a contact
\end{description}

\subsubsection{Score Types}

\subsubsection{The Heist}

Infiltrate a secure location to steal valuable items or information.

\textbf{Key Clocks:}
\begin{itemize}
    \item Alarm [4-6] - Security response time
    \item Security Sweep [4] - Methodical search patterns
    \item Extraction Window [4] - Time-limited escape opportunity
\end{itemize}

\textbf{Approach Vectors:}
\begin{itemize}
    \item \textbf{Deceit} - Social engineering and false identities (Presence + Sway/Deception)
    \item \textbf{Stealth} - Avoiding detection through shadows and silence (Body + Stealth)
    \item \textbf{Violence} - Direct confrontation when subtlety fails (Body + Melee/Athletics)
    \item \textbf{Bribery} - Paying for access or silence (Wits + Streetwise)
\end{itemize}

\subsubsection{The Smuggling Run}

Transport contraband through checkpoints and inspections.

\textbf{Key Clocks:}
\begin{itemize}
    \item Papers Scrutiny [6] - Document examination intensity
    \item Patrol Net [6] - Law enforcement presence
    \item Cargo Integrity [4] - Condition of smuggled goods
\end{itemize}

\textbf{Special Procedures:}
\begin{itemize}
    \item \textbf{Hidden Compartments:} Smuggle DV -1; on 1s, start Contraband Trail [4]
    \item \textbf{Cutout Operations:} Delegate checkpoint interactions to reduce personal exposure
    \item \textbf{Night Passage:} Stealth approaches during low-visibility conditions
\end{itemize}

\subsubsection{The Protection Play}

Defend a person, place, or operation from rivals or authorities.

\textbf{Key Clocks:}
\begin{itemize}
    \item Threat Assessment [4] - Intelligence about incoming dangers
    \item Security Posture [6] - Defensive preparations
    \item Response Time [4] - Speed of reaction to threats
\end{itemize}

\textbf{Defensive Options:}
\begin{itemize}
    \item \textbf{Safe Houses:} Ghost Safehouses (String) for emergency refuge
    \item \textbf{Lookouts:} Early warning systems using local contacts
    \item \textbf{Counter-Surveillance:} Active measures to detect and evade watchers
\end{itemize}

\subsubsection{The Confidence Game}

Manipulate targets through elaborate schemes and false trust.

\textbf{Key Clocks:}
\begin{itemize}
    \item Trust Building [6] - Establishing credibility with marks
    \item Scheme Complexity [4] - Maintaining multiple deception layers
    \item Exposure Risk [6] - Chance of being discovered
\end{itemize}

\textbf{Social Engineering Tools:}
\begin{itemize}
    \item \textbf{Cover Identities:} False backgrounds with supporting documentation
    \item \textbf{Emotional Manipulation:} Exploiting targets' desires and fears
    \item \textbf{Evidence Plants:} Creating false proof to support deceptions
\end{itemize}

\subsection{Heat Management During Heists}

\subsubsection{Setup Rule}

Each SB gained during Setup may either:
\begin{itemize}
    \item Alarm +1 (advance security response)
    \item Crew Heat +1 (increase personal attention)
\end{itemize}
GM chooses based on fiction.

\subsubsection{Heat Events During Scores}

\textbf{Crew Heat 3:} Patrol Sweep [4] begins
\textbf{Crew Heat 6:} Trigger Sting [6] and Exposure +1
\textbf{City Heat 3:} Random stops; Notoriety +1 when making flashy moves
\textbf{City Heat 6:} Crackdown [4] (DV +1 on criminal actions)

\subsubsection{Cooling Actions}

\begin{description}
    \item[Lay Low] Lose 1 Bandwidth; clear Crew Heat -2; gain Notoriety +1
    \item[Grease Palms] Spend Bank/Favor; clear Crew Heat -1 and one Inspection
    \item[Hard Pivot] Burn a front to clear Crew Heat -3 and City Heat -1
\end{description}

\subsubsection{Informant Networks in Heists}

\subsubsection{Cross-Check Advantage}

Spend 2 informants on the same topic:
\begin{itemize}
    \item Take +1 Effect on related rolls
    \item Only the lowest Reliability -1 applies on a 1
\end{itemize}

\subsubsection{Masking Sources}

Spend Bank 1 to prevent Reliability -1 once when using informants.

\subsubsection{Informant Complications}

When an informant's Reliability reaches 0:
\begin{itemize}
    \item They flip allegiance or vanish
    \item Start Burn [4] to recover/replace
    \item GM may spend SB to create immediate problems
\end{itemize}

\subsection{Black Market Integration}

\subsubsection{Acquiring Heist Gear}

Market Tiers and DV:
\begin{itemize}
    \item Street DV 2-3
    \item Guild DV 3-4  
    \item Court DV 4-5
    \item Arcane DV 5 (wards, psionic keys, oath-lamps)
\end{itemize}

\subsubsection{Partials and Misses}

\textbf{Partials:}
\begin{itemize}
    \item Price spike (Bank -1)
    \item Crew Heat +1
    \item Quality flaw (-1 Effect next use)
    \item Seller's Temp String
    \item Inspection [4] starts
\end{itemize}

\textbf{Misses:}
\begin{itemize}
    \item Sting springs (Sting [6] +2)
    \item Counterfeit gear
    \item Owed favor to rival fence
\end{itemize}

\subsubsection{Heist Complications Generator}

\begin{center}
\begin{tabular}{|c|p{10cm}|}
\hline
\textbf{Roll 1d12} & \textbf{Complication} \\
\hline
1 & Inside man panics \\
2 & Boom winch jammed; requires at-sea repair under watchfire \\
3 & Spirit net across eel-cut; needs Ward or sacrifice \\
4 & Lantern tower replaced crew; new officer loves procedure \\
5 & Cargo stinks of contraband spice; dogs alert \\
6 & Friendly convoy is bait; hidden interdictor behind fog \\
7 & Rival smugglers mirror your codes; collisions of story \\
8 & Weather window slams; cross chop makes boarding Dominant \\
9 & Ghost ballast sings; occult backwash risks crew morale \\
10 & Courtesy duel demanded by a junior; public \\
11 & Harbormaster's nephew aboard; scandal risk \\
12 & Prisoner knows you; shouts your alias \\
\hline
\end{tabular}
\end{center}

\subsubsection{Heist Rewards and Consequences}

\subsubsection{Successful Heist Outcomes}

\begin{description}
    \item[Clean Score] 3-5 Cred + Standing with criminal elements
    \item[Heat Management] Reduce Crew Heat -1 through careful planning
    \item[New Contacts] Gain informants or fence relationships
    \item[Reputation Building] Increase Notoriety in appropriate circles
\end{description}

\subsubsection{Failed Heist Consequences}

\begin{description}
    \item[Sting Operations] Trigger law enforcement responses
    \item[Exposure] Increase personal and crew visibility
    \item[Debt Accumulation] Owe favors to fix problems
    \item[Reputation Damage] Lose Standing with criminal networks
\end{description}

\subsubsection{Heist Scene Examples}

\subsubsection{The Vault Break}

\textbf{Setup:} Infiltrate merchant's house to access secure safe

\textbf{Position:} Controlled (inside contact), DV 3

\textbf{Key Actions:}
\begin{itemize}
    \item Disable (Wits + Tinker) vs. DV 4 for lock mechanisms
    \item Infiltrate (Body + Stealth) vs. DV 3 for guard avoidance
    \item Skulk/Shadow (Wits + Subterfuge) vs. DV 2 for timing
\end{itemize}

\textbf{Threat Clocks:}
\begin{itemize}
    \item House Guards [6]
    \item Vault Mechanism [4] 
    \item Alarm Response [4]
\end{itemize}

\subsubsection{The Social Heist}

\textbf{Setup:} Infiltrate noble's ball to steal documents

\textbf{Position:} Risky (under scrutiny), DV 3

\textbf{Key Actions:}
\begin{itemize}
    \item Sway (Presence + Sway) vs. DV 3 for social access
    \item Deception (Presence + Deception) vs. DV 4 for false identity
    \item Notice (Wits + Notice) vs. DV 2 for security observation
\end{itemize}

\textbf{Threat Clocks:}
\begin{itemize}
    \item Guest Recognition [6]
    \item Security Patrols [4]
    \item Document Location [4]
\end{itemize}

\subsubsection{Heist Scaling Guidelines}

\subsubsection{Small Crew (2-3 members)}

\begin{itemize}
    \item Focus on 1-2 approach vectors
    \item Reduce clock sizes by 1-2 segments
    \item Increase DV by 1 for complex operations
    \item Emphasize individual skill over team coordination
\end{itemize}

\subsubsection{Large Crew (4-6 members)}

\begin{itemize}
    \item Multiple simultaneous approach vectors
    \item Increase clock sizes by 1-2 segments
    \item Decrease DV by 1 for coordinated efforts
    \item Emphasize teamwork and specialization
\end{itemize}

\subsubsection{High-Stakes Operations}

\begin{itemize}
    \item Increase all DV by 1-2
    \item Add additional threat clocks
    \item Reduce Heat management options
    \item Increase consequences for failure
\end{itemize}

\subsubsection{Heist Integration with Other Systems}

\subsubsection{Political Intrigue}

\begin{itemize}
    \item Heists can gather intelligence for larger schemes
    \item Political cover can reduce Heat consequences
    \item Faction relationships affect market access
    \item Court connections provide legal cleanup options
\end{itemize}

\subsubsection{Assets and Patrons}

\begin{itemize}
    \item Safe houses provide extraction points
    \item Patron influence can reduce legal consequences
    \item Asset leverage creates additional approach vectors
    \item Strings facilitate market access for stolen goods
\end{itemize}

\subsubsection{Allies and Adversaries}

\begin{itemize}
    \item Followers can provide specialized assistance
    \item Rival crews create competitive pressure
    \item Informant networks enable better planning
    \item Organizational backing provides resources
\end{itemize}

\section{Tools for Play}
\subsection{Typical Actions (DV)}
Infiltrate 2–4 • Disable 3–5 • Skulk/Shadow 2–4 • Broker 2–4 • Intimidate/Break 2–4 • Evade 2–4 • Operate (Smuggle) 2–4.

\subsection{Chases (Urban Ladder)}
Foot: 2 open • 3 markets • 4 rooftops • 5 sewers. \\
Cart/Boat: 2 canal • 3 locks • 4 eel‑cuts • 5 tide race.

\subsection{Heat Events (thresholds)}
At Crew Heat 3: Patrol Sweep [4]. \\
At Crew Heat 6: Sting [6] + Exposure +1. \\
At City Heat 3: random stops/curfew murmurs. \\
At City Heat 6: Crackdown [4] (DV +1).

\subsection{Shadows \& Steel SB (GM menu)}
Marked Silver (Inspection [4]) • Bad Badge (Alarm +1) • Gossip Sting (Audience flips) • Dogged Tail (Pursuit [6]) • Union Pushback (Blood‑Silver owed) • Turned Informant (Reliability –1) • Watch List (Crew Heat +1).

\subsection{Overhead Aids (NEW)}
\begin{itemize}
\item \textbf{Clock Budget (3+1):} keep three visible clocks (Objective, Alarm/Patrol, a Rival/War) + one shadow (GM). Fold extra pressures into SB spends.
\item \textbf{Heat→Complication:} when tempted to add a new clock mid‑score, raise Crew Heat +1 and spend an SB item instead.
\end{itemize}

\section{Generators}
\subsection*{Black Market Goods (d12):} 1 rare pigment • 2 forged seal kit • 3 ward key • 4 eel‑channel charts • 5 customs stamps • 6 moon‑kiln glass • 7 lockspike set • 8 poison salts • 9 tax records • 10 theater takings • 11 relic case • 12 ledger cipher.

\subsection*{Fences \& Fixers (d12):} 1 archivolt broker • 2 quay factor • 3 lamp‑notary • 4 skald bookie • 5 guild widow • 6 patrol sergeant • 7 foreign factor • 8 tie‑in priest • 9 river witch • 10 street doctor • 11 dice lord • 12 mask troupe.

\subsection*{Law Gambits (d12):} 1 bait coin • 2 uniform sweep • 3 warrant blitz • 4 lamplighter eaves • 5 snitch buy • 6 courier sting • 7 tax audit • 8 curfew test • 9 postern watch • 10 vessel inspection • 11 customs raid • 12 proof‑of‑life stall.

\section{Example of Play (short)}
As in v0.1, updated to reflect Setup SB → Heat/Alarm and Clock Budget. (See canvas appendix for the full "Dyer's Ledger" walkthrough.)

\section{Long Arcs (NEW)}
\begin{itemize}
\item \textbf{Ledger War:} expose or control citywide tax records; milestones: seize weighhouse seal • flip a lamp‑notary • bargain with Tax Assessor • settle with guilds.
\item \textbf{Convoy Black:} dominate river/port smuggling; milestones: eel‑channel map • union détente • patrol captain's debt • free harbor writ.
\item \textbf{Ward Grid:} steal/broker arcane infrastructure; milestones: ward‑glass salvage • oath‑lamp control • ruin‑keeper pact • mirrored hall heist.
\end{itemize}

\section{Political Integration (NEW)}
Link War [8] to Political Faction Clocks: each War +2 also ticks a related Influence/Stability/Exposure +1 for the two factions most involved. Settlement with witness faction grants Favor +1 with that faction if the oath stands a season.

\section{GM Reference (one page)}
\begin{itemize}
\item Crew: tags, Strings, Crew Heat [6], Notoriety (opt.), Rep.
\item Turf Map: nodes; Claim [6]; War [8]; settlement template.
\item Heat: thresholds; Sweep → Sting → Crackdown; Cooling.
\item Markets: tiers (Street/Guild/Court/Arcane); partials; laundering.
\item Informants: Web; Reliability [4]; Cross‑Check/Mask.
\item Scores: planner; Kit Tokens; threat clocks.
\item SB Menu \& Overhead: SB list; Clock Budget (3+1); Heat→Complication.
\item Integration: V\&Stone districts; Political factions; Caravan smuggling; psionic/ward crimes.
\end{itemize}

\section*{Changelog}
v0.2 — Added Fixers \& Ward‑Breakers; formal Heat thresholds and Setup rule; Arcane Tier market; supernatural crimes; overhead aids; long arcs; political mapping; settlement template; revised example pointers.

End of v0.2

\end{document}