\documentclass[12pt,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathpazo}
\usepackage{lettrine}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{epigraph}
\usepackage{verse}
\usepackage{multicol}
\usepackage{titlesec}
\usepackage[margin=1in]{geometry}
\usepackage{afterpage}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}

% Page styling
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[LO]{\leftmark}
\fancyhead[RE]{\rightmark}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% Chapter styling
\titleformat{\chapter}[display]
{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}
\titlespacing*{\chapter}{0pt}{50pt}{40pt}

% Custom commands
\newcommand{\ritualname}[1]{\vspace{10pt}\noindent\textbf{\large #1}\vspace{5pt}}
\newcommand{\shadow}[1]{\textit{#1}}
\newcommand{\whisper}[1]{\textbf{#1}}

% PRODUCT TITLE
% Use on cover, title page, and store listings
\newcommand{\ExpansionTitle}{Fate's Edge: The Whispered Grimoire}
\newcommand{\ExpansionSubtitle}{Rituals \& Rites}
% Requires \usepackage{hyperref}
% SHORT INTERNAL HANDLE (headers, refs)
\newcommand{\ShortRunekeeper}{The Whispered Grimoire}

\begin{document}

\frontmatter

% Title Page
\begin{titlepage}
\centering
\vspace*{2cm}
{\Huge\bfseries Fate's Edge: The Whispered Grimoire\par}
\vspace{1cm}
{\Large A Collection of Rituals, Rites and Forbidden Arts\par}
\vspace{2cm}
{\Large Compiled by Veyla the Hedge-Witch\par}
\vspace{0.5cm}
{\large Servant of Ikasha, She Who Sleeps Within Shadow\par}
\vspace{1cm}
{\large With Notes from Rukhir, My Raven Familiar\par}
\vspace{3cm}
{\large Tenth Moon of the Crimson Year\par}
\vspace{0.5cm}
{\large From the Press of Whispered Pages, Theona\par}
\end{titlepage}

% Copyright Page
\newpage
\thispagestyle{empty}
\vspace*{5cm}
\begin{center}
{\small
The Whispered Grimoire\\
A Collection of Shadow-Rituals and Forbidden Arts\\
\vspace{0.5cm}
Compiled by Veyla the Hedge-Witch\\
\vspace{1cm}
Copyright \copyright\ 2025 Veyla of the Tulkani\\
All rights reserved. No part of this work may be reproduced or transmitted in any form or by any means without written permission from the author.\\
\vspace{1cm}
This work contains dangerous knowledge. Read at your own risk.\\
The author and publisher are not responsible for any consequences arising from the use of these rituals.\\
\vspace{1cm}
Printed in the Three Greens\\
Press of Whispered Pages, Theona\\
Tenth Moon of the Crimson Year
}
\end{center}

% Dedication
\newpage
\thispagestyle{empty}
\vspace*{8cm}
\begin{center}
\begin{verse}
\textit{To Rukhir, my shadow-brother,}\\
\textit{whose black wings carry secrets} \\
\textit{and whose beak pecks at truth} \\
\textit{until it bleeds revelation.} \\
\ \\
\textit{To Ikasha, Mistress of Whispers,}\\
\textit{who teaches us that in shadow} \\
\textit{we find the spaces between} \\
\textit{where power truly dwells.}
\end{verse}
\end{center}

% Table of Contents
\tableofcontents

\mainmatter

\chapter{To the Seeker of Shadows}

\lettrine[lines=3]{L}{}et me begin, dear reader, with a warning whispered in the dark hours before dawn, when the veil between worlds grows thin as a spider's silk. What you hold in your hands is not mere parchment and ink, but a gathering of truths that the light-bound would burn, the law-bound would imprison, and the truth-bound would deny.

I am Veyla, hedge-witch of the Tulkani, daughter of wandering merchants who taught me that the greatest treasures are those hidden in plain sight. My path led me to the worship of Ikasha, She Who Sleeps Within Shadow, and through her dark blessings, I have learned that power flows not from the sun's harsh light, but from the spaces between the stars.

\shadow{Rukhir tells me you are worthy of these words. He has been watching you, as ravens watch all who walk the paths between. "She moves with the caution of one who knows that shadows have teeth," he caws to me. "Her footsteps make no sound on the stones of secrecy." For a raven to speak of caution is high praise indeed.}

The rituals contained within these pages are not for the faint of heart or the weak of spirit. Each one demands a price, for Ikasha's gifts are never given freely. But for those with the wit to pay and the will to claim, they offer power that kings would kill for and gods would envy.

\section*{The Nature of Shadow-Magic}

Shadow is not darkness, child of light. It is the space between what is and what could be, the pause between heartbeats where truth whispers its name. In my years of walking the twilight paths, I have learned that shadow-magic is not about casting spells in the traditional sense, but about opening doors that others have locked and forgotten.

\shadow{Rukhir nods from his perch upon my shoulder. "The simple think shadow is absence," he says, preening his black feathers. "The wise know it is presence waiting to be acknowledged."}

The Tulkani understand this better than most peoples. We are wanderers, yes, but we are also watchers. In our travels, we have learned that every culture hides its secrets in the spaces between words, and it is in those spaces that true power lies. The shadow-magics I present here are drawn from a dozen traditions, whispered to me by spirits, traded with other seekers, and discovered through my own communion with Ikasha.

\section*{The Price of Power}

Before we begin, you must understand a fundamental truth: there is no such thing as free magic. The Utarans speak of balance, the Aeler of the deep laws, and the Wood-Elves of the cycle of giving and taking. In the tongue of my people, we say that every gift must be paid for in kind, and Ikasha's gifts are the most expensive of all.

\whisper{Listen well to this, for it is the most important lesson I can teach you:}

\begin{verse}
The shadow that you command \\
Will command you in return. \\
The secret that you steal \\
Will steal a piece of you. \\
The truth that you uncover \\
Will uncover truths about yourself \\
That you may not wish to know.
\end{verse}

\shadow{Rukhir caws agreement. "I have seen wizards become slaves to their own spells," he tells me. "I have watched sorceresses consumed by the very shadows they thought to control."}

This is why I have included not just the rituals themselves, but the warnings, the preparations, and the aftermaths. Each ritual in this grimoire comes with three sections: The Weaving (how to perform the ritual), The Price (what you must give), and The Whisper (what Ikasha will whisper to you in return).

\section*{The Tools of the Shadow-Witch}

You will find that shadow-magic requires few material components compared to the flashy spells of the light-wielders. Our power comes from understanding rather than force, from knowing rather than doing. However, there are certain tools that every shadow-witch should possess:

\begin{itemize}
\item \textbf{A black candle} - Not for light, but for focus. The flame should gutter and dance as if in a wind that others cannot feel.
\item \textbf{A mirror of dark glass} - To see what others cannot see, and to guard against what would see you.
\item \textbf{A raven's feather} - For writing in shadow-ink that only the initiated can read.
\item \textbf{A small pouch of shadow-salt} - Gathered at the crossroads when the moon is new, used to mark boundaries and ward off unwanted attention.
\item \textbf{A whisper-stone} - A smooth stone that has been held in the mouth during a silence, used to store and release spoken secrets.
\end{itemize}

\shadow{Rukhir has brought me a bone from a hanged man, which he claims is essential for certain workings. I have not yet had the courage to use it.}

\section*{The Structure of This Work}

This grimoire is divided into four sections, each corresponding to a different aspect of shadow-magic:

\begin{enumerate}
\item \textbf{Whispers and Secrets} - Rituals for gathering information, speaking across distances, and uncovering hidden truths.
\item \textbf{Shadows and Veils} - Rituals for concealment, transformation, and movement through spaces that should not exist.
\item \textbf{Dreams and Nightmares} - Rituals for entering the dream-realm, influencing sleepers, and drawing power from the unconscious mind.
\item \textbf{Bargains and Bindings} - Rituals for making pacts with spirits, binding servants, and compelling others to your will.
\end{enumerate}

Each ritual is presented in the same format:

\ritualname{Name of the Ritual}
\textbf{Purpose:} What the ritual accomplishes\\
\textbf{Weaving:} Step-by-step instructions for performance\\
\textbf{Price:} What must be sacrificed or risked\\
\textbf{Whisper:} What Ikasha grants in return\\
\textbf{Warning:} Important cautions and potential consequences

\shadow{Rukhir insists I add this final note: "Not all who seek power are worthy of it, and not all who are worthy seek it. Be certain which you are before you turn the next page."}

Now, if you have read this far and still wish to continue, then perhaps you are ready for what lies ahead. Light your black candle, settle into your chair, and prepare to learn the arts that others fear to name.

But remember, dear seeker - once you have seen into the shadows, you can never claim to walk in the light again.

\chapter{Whispers and Secrets}

\section*{The Art of Listening}

In the Tulkani tradition, we say that the greatest magic is not in speaking, but in hearing what others dare not say. The rituals in this chapter will teach you to listen to the whispers that exist in the spaces between words, to read the truth in a person's eyes, and to gather secrets that would make emperors tremble.

\shadow{Rukhir has been very quiet this evening. I think he senses that we are approaching dangerous territory. Even ravens have their secrets, it seems.}

\ritualname{The Listening Shadow}
\textbf{Purpose:} To hear conversations from great distances, particularly those that the speakers believe are private.

\textbf{Weaving:} 
\begin{enumerate}
\item Light a black candle at the darkest hour of night, when even the moon has set.
\item Take a raven's feather and whisper into it the name of the person whose words you wish to hear.
\item Bury the feather beneath your doorstep, or in the earth beneath a window where your target might pass.
\item For three nights, you must dream of shadows dancing. Do not fight these dreams - embrace them.
\item On the fourth night, the words you seek will come to you on the wind, spoken in a voice like rustling leaves.
\end{enumerate}

\textbf{Price:} You will hear one conversation that you desperately wish not to hear. This knowledge will burden you until you either act upon it or deliberately forget it.

\textbf{Whisper:} Ikasha will grant you the ability to distinguish truth from lies in any spoken word, but only for as long as you hold your breath.

\textbf{Warning:} Do not use this ritual to spy upon those who are innocent of wrongdoing. Ikasha favors those who seek justice, not those who indulge in cruelty.

\shadow{Rukhir shifts uncomfortably. "I have seen this ritual drive people to madness," he mutters. "The knowledge comes with the weight of responsibility."}

\section*{The Bonding Ceremony}
\textit{A ritual to strengthen connections between companions}

\textbf{Purpose:} Create a permanent bond between party members that provides mechanical and narrative benefits

\textbf{Participants:} All willing party members (3-5 helpers maximum)

\textbf{Channel Phase (Wits + Bonds):}
\begin{enumerate}
\item Each participant holds an object representing their connection to the group
\item Perform a shared activity (sharing a meal, telling stories, etc.) while focusing on the bond
\item Roll Wits + Bonds skill for each participant
\end{enumerate}

\textbf{Weave Phase (Presence + Sway):}
\begin{enumerate}
\item Participants speak words of commitment to each other
\item Create a physical symbol of the bond (knotting cords, mixing earth from homelands, etc.)
\item Roll Presence + Sway for the primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} Temporary tension between bonded members (-1 die to cooperative actions for Significant Time)
\item \textbf{3-4 CP:} The bond creates unwanted attention from external forces; party generates 2 CP whenever they act as a group for the next session
\item \textbf{5+ CP:} All existing bonds between party members are temporarily severed; party suffers -2 dice to all social rolls for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Bonded members gain +1 die when aiding each other for one arc
\item Can sense each other's general emotional state across any distance for one arc
\item One "bond-mate" can be called upon for aid once per session (counts as Cap 3 follower for that action)
\end{itemize}

\shadow{Rukhir preens thoughtfully. "I have seen flocks torn apart by too-strong bonds," he caws. "The heart that beats as one can also break as one."}

\textbf{Duration:} One full arc, renewable with simpler weekly reaffirmation

\textbf{Casting Time:} Significant Time (ritual requires shared meal and ceremony)

\section*{The Sanctuary Consecration}
\textit{A ritual to create a safe haven from supernatural threats}

\textbf{Purpose:} Establish a protected area that repels hostile supernatural entities

\textbf{Primary Caster:} Spirit 4+, Lore 3+

\textbf{Helpers:} 2-3 participants with relevant skills (Lore, Faith, Arcana)

\textbf{Channel Phase (Spirit + Faith/Lore):}
\begin{enumerate}
\item Gather materials sacred to the caster's beliefs or relevant to the threats faced
\item Consecrate the space through prayer, meditation, or ritual preparation
\item Roll Spirit + relevant skill for each participant
\end{enumerate}

\textbf{Weave Phase (Wits + Ritual):}
\begin{enumerate}
\item Perform the consecration ceremony within the designated area
\item Create permanent wards using blessed/sacred materials
\item Roll Wits + Ritual for the primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The sanctuary attracts the attention of other supernatural entities; 2 CP added to next supernatural encounter
\item \textbf{3-4 CP:} The protection creates a vulnerability elsewhere; choose one party member who becomes more susceptible to supernatural influence for one session
\item \textbf{5+ CP:} The ritual backfires, making the area supernaturally dangerous; the space becomes a beacon for hostile entities
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Area becomes sanctified ground (Sanctity +2) against specified supernatural threats for one month
\item Hostile entities of the specified type suffer -2 dice to actions within the area
\item Area provides +1 die to resistance rolls against specified supernatural effects for one month
\item Party members resting within the sanctuary recover 1 additional Fatigue per night
\end{itemize}

\shadow{Rukhir circles the perimeter warily. "Sacred spaces are like nests," he observes. "They keep some things out, but they also keep you in."}

\textbf{Duration:} One month, renewable with weekly maintenance (10 minutes of ritual work)

\textbf{Casting Time:} Significant Time (requires preparation and ceremony)

\section*{The Memory Reconstruction}
\textit{A ritual to recover lost or suppressed memories}

\textbf{Purpose:} Retrieve information from a participant's suppressed or forgotten memories

\textbf{Primary Caster:} Wits 4+, Arcana 3+

\textbf{Helpers:} 1-2 participants with Insight or Lore skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Participant enters a deep meditative state
\item Primary caster prepares scrying tools or ritual components
\item Roll Wits + Arcana for primary caster, Wits + Insight/Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Spirit + Insight):}
\begin{enumerate}
\item Guide the participant through their memories using guided visualization
\item Use ritual tools to stabilize recovered memories
\item Roll Spirit + Insight for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} Recovered memories are fragmented or unreliable; information may be misleading
\item \textbf{3-4 CP:} The ritual triggers traumatic memories causing temporary psychological effects; participant suffers -1 die to social rolls for Significant Time
\item \textbf{5+ CP:} Participant becomes obsessed with recovered memories; must make Spirit + Resolve (DV 2) rolls to act against the memory's implications for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Recover one specific memory or set of memories from the participant's past (immediate)
\item Memory clarity depends on success margin (partial recall vs. complete recovery)
\item Participant gains +2 dice on related knowledge rolls for this scene
\item GM reveals one piece of information that connects the recovered memory to current events
\end{itemize}

\shadow{Rukhir tilts his head. "The past is a cage made of whispers," he says. "Some birds prefer to forget they were ever caged."}

\textbf{Duration:} Memory effects last for this scene; knowledge benefits last for this session

\textbf{Casting Time:} Some Time (requires deep meditation and guided ritual)

\section*{The Shadow Passage}
\textit{A ritual to travel through the Ways Between}

\textbf{Purpose:} Transport the party across great distances through shadow-realm shortcuts

\textbf{Primary Caster:} Wits 5+, Arcana 4+, Spirit 3+

\textbf{Helpers:} All party members (spiritual support during the journey)

\textbf{Channel Phase (Spirit + Arcana):}
\begin{enumerate}
\item Create a circle of shadow-conductive materials (obsidian, black candles, etc.)
\item Each participant contributes a drop of blood or lock of hair to the circle
\item Roll Spirit + Arcana for primary caster, Spirit + Resolve for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Ritual):}
\begin{enumerate}
\item Perform the opening ceremony to breach the shadow-realm
\item Lead the party through the passage while maintaining the connection
\item Roll Wits + Ritual for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The journey takes longer than expected, arriving at an inconvenient time (advance relevant clocks by 1 segment)
\item \textbf{3-4 CP:} The party arrives in a slightly different location than intended; position for first action is Risky instead of Controlled
\item \textbf{5-6 CP:} Something follows the party back from the shadow-realm; introduce a persistent supernatural threat that tracks the party
\item \textbf{7+ CP:} The party finds themselves in the Ways Between.
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Transport party instantly across great distances (anywhere on the same continent) for this scene
\item Arrival position is Controlled for planned destinations, Risky for improvised ones
\item +1 effect on the first action after arrival (momentum from the journey)
\item Party is immune to fatigue from the travel itself
\item Cannot be used more than once per week without increasing backlash severity
\end{itemize}

\shadow{Rukhir fluffs his feathers nervously. "The spaces between spaces have their own hunger," he warns. "They take payment in ways you won't expect until it's too late."}

\textbf{Duration:} Instantaneous travel with lingering effects for this scene

\textbf{Casting Time:} Significant Time (complex ritual requiring careful preparation)

\section*{The Geas of Binding}
\textit{A ritual to create unbreakable oaths or compulsions}

\textbf{Purpose:} Bind a willing or unwilling target to a specific course of action

\textbf{Primary Caster:} Presence 4+, Spirit 4+, Lore 3+

\textbf{Helpers:} 1-2 participants with relevant cultural or religious knowledge

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Prepare the binding materials according to cultural tradition
\item Research the target's nature and vulnerabilities
\item Roll Spirit + Lore for primary caster, relevant skill for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Perform the binding ceremony with the target present
\item Speak the terms of the geas in the appropriate ritual language
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The geas has unintended loopholes or consequences; target finds creative ways to technically fulfill the geas while undermining its intent
\item \textbf{3-4 CP:} The caster becomes partially bound by the same geas; must also follow a related constraint for Significant Time
\item \textbf{5+ CP:} The binding creates a spiritual debt that attracts unwanted supernatural attention; 3 CP added to next supernatural encounter involving the caster
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Target is compelled to follow the terms of the geas or suffer spiritual consequences for one arc
\item Breaking the geas requires a successful Spirit + Resolve roll (DV 4+) each time
\item Failure to follow the geas causes 1 CP per day of violation
\item Geas can be complex but must be clearly defined (no "be good" - specify actions)
\item Willing targets suffer reduced backlash; unwilling targets may resist with opposed roll
\end{itemize}

\shadow{Rukhir caws sharply. "Words are chains made of air," he says. "But they bind tighter than iron when spoken with true power."}

\textbf{Duration:} One full arc, permanent unless specifically countered by equal magic

\textbf{Casting Time:} Significant Time (formal ceremony requiring proper ritual components)

\section*{The Communion of the Ancients}
\textit{A ritual to speak with the spirits of the dead}

\textbf{Purpose:} Contact and communicate with deceased individuals or ancestral spirits

\textbf{Primary Caster:} Spirit 4+, Lore 3+, Faith 2+

\textbf{Helpers:} 1-2 participants with spiritual sensitivity or relevant cultural knowledge

\textbf{Channel Phase (Spirit + Faith):}
\begin{enumerate}
\item Prepare the ritual space according to cultural traditions for contacting the dead
\item Offer appropriate sacrifices or prayers to the spirits
\item Roll Spirit + Faith for primary caster, Spirit + Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Lore):}
\begin{enumerate}
\item Open the channel to the spirit realm
\item Guide the summoned spirit toward communication
\item Roll Wits + Lore for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The contacted spirit is confused or unhelpful; information provided is vague or misleading
\item \textbf{3-4 CP:} Other, less welcome spirits are drawn to the ritual; 2 CP added to next supernatural encounter
\item \textbf{5+ CP:} The ritual creates a permanent spiritual connection that cannot be easily severed; caster suffers -1 die to all social rolls involving living beings for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Communicate with one specific deceased individual or type of ancestral spirit for Some Time
\item Spirit can provide information, advice, or warnings about specified topics
\item Spirit's knowledge is limited by their mortal understanding and perspective
\item Communication lasts for a number of minutes equal to the success margin
\item Spirit may request a favor or service in return for their cooperation
\end{itemize}

\shadow{Rukhir shivers, feathers ruffling. "The dead remember everything," he whispers. "But they forget what it means to be kind to the living."}

\textbf{Duration:} Some Time (ritual communication period)

\textbf{Casting Time:} Significant Time (preparation and ceremony required)

\section*{The Ward of Protection}
\textit{A ritual to create lasting magical defenses}

\textbf{Purpose:} Establish permanent or semi-permanent protective wards

\textbf{Primary Caster:} Wits 4+, Arcana 3+, relevant cultural skill

\textbf{Helpers:} 2-3 participants with Craft, Engineering, or cultural ritual skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Gather materials appropriate to the type of protection desired
\item Prepare the location where wards will be placed
\item Roll Wits + Arcana for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Craft/Engineering):}
\begin{enumerate}
\item Inscribe, construct, or otherwise create the physical components of the wards
\item Activate the wards through ritual consecration
\item Roll Wits + relevant skill for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The wards have a minor vulnerability or require specific activation conditions; wards function at -1 effectiveness until properly calibrated
\item \textbf{3-4 CP:} The wards attract the attention of entities they're designed to repel; 2 CP added to next encounter with specified threat type
\item \textbf{5+ CP:} The wards backfire, creating the opposite effect or affecting unintended targets; wards harm allies instead of protecting them for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create wards that provide +2 dice to defense against specified threats for one month
\item Wards can cover an area, protect specific individuals, or guard objects
\item Duration: One month, renewable with monthly maintenance rituals
\item Wards can be complex (multiple effects) but each additional effect increases backlash risk
\item Wards can be detected by those with appropriate skills and may reveal the caster's presence
\end{itemize}

\shadow{Rukhir examines the warding materials critically. "Boundaries are promises written in power," he observes. "But promises can be broken, and power can be hungry."}

\textbf{Duration:} One month, renewable with weekly maintenance

\textbf{Casting Time:} Significant Time (complex construction and consecration process)

\section*{The Veil of Misdirection}
\textit{A ritual to conceal the party's true nature and intentions}

\textbf{Purpose:} Create illusions and false impressions that mask the party's activities

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Deception 2+

\textbf{Helpers:} 2-3 participants with Performance, Deception, or cultural masking skills

\textbf{Channel Phase (Wits + Deception):}
\begin{enumerate}
\item Prepare illusion components (cosmetics, costumes, symbolic objects)
\item Each participant contributes a personal item that represents their true self
\item Roll Wits + Deception for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Performance):}
\begin{enumerate}
\item Perform the masquerade ritual, weaving false identities around each participant
\item Create shared illusion patterns that reinforce the group's fabricated story
\item Roll Presence + Performance for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The illusions have minor flaws or inconsistencies; -1 die to social rolls when under scrutiny
\item \textbf{3-4 CP:} The false identities begin to influence the participants' behavior; party members must make Wits + Resolve (DV 2) rolls to act against their assumed personas for Significant Time
\item \textbf{5+ CP:} The ritual creates a feedback loop where the false identities become partially real; participants suffer identity confusion and -2 dice to rolls involving their true skills for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Party assumes completely fabricated identities that are believable to casual observers for one session
\item Each participant gains +2 dice to Deception and Performance rolls related to their assumed identity
\item Social interactions start Controlled when the false identities are relevant
\item Magical detection of true identities suffers -2 dice for Significant Time
\item Party can coordinate complex deceptions as a group action with +1 effect
\end{itemize}

\shadow{Rukhir preens his feathers with satisfaction. "I have worn many faces," he crows. "But I always remember which one is mine."}

\textbf{Duration:} One full session, requiring daily renewal (10 minutes of ritual work)

\textbf{Casting Time:} Significant Time (elaborate preparation and group ceremony)

\chapter{Part II: The Shadow's Embrace}

\section*{On Possession and the Compromised Soul}

\lettrine[lines=3]{T}{}he deepest shadows, dear reader, are not those that fall from candlelight or moonbeam. They are the darkness that grows within, fed by our own fears, our own compromises, our own willingness to trade pieces of ourselves for power. 

\shadow{Rukhir has not spoken much of late. I think he senses the weight of what I must record in these pages. Even ravens, I have learned, can be touched by shadows that should remain buried.}

In all my years walking the twilight paths, I have seen many seekers of power become its prisoners. The line between master and servant is thinner than spider's silk, and it snaps more easily than most would believe. When I speak of possession, I do not merely mean the crude taking of bodies by wandering spirits—though that too has its place in the dark arts. No, the possession I warn you of is subtler, more insidious, and far more dangerous.

\textbf{Player Character Possession:} This is not a condition to be taken lightly, nor one that should ever be inflicted without the player's consent and understanding. Possession represents the ultimate compromise of character agency—the point at which external forces have so thoroughly corrupted or influenced a character that they are no longer fully in control of their actions.

The Compromised condition of possession manifests in stages:

\begin{itemize}
\item \textbf{Stage 1 (Influenced):} Character begins to hear whispers, feel urges, or experience compulsions that are not their own. They retain full agency but must resist temptation through appropriate rolls.
\item \textbf{Stage 2 (Controlled):} Character occasionally acts against their will or better judgment. The possessing force can compel specific actions once per session, which the player may resist with a Spirit + Resolve roll.
\item \textbf{Stage 3 (Possessed):} Character becomes a puppet of the possessing entity. The GM takes control of the character for extended periods, though major life-or-death decisions should still involve player consultation.
\end{itemize}

\shadow{"Possession is not defeat," Rukhir finally speaks, his voice unusually grave. "It is transformation. And transformation, my dear Veyla, is always a kind of death."}

\textbf{GM Guidance on Possession:}

Possession should never be used as a punitive measure or to railroad players. It is a narrative tool of last resort, representing the culmination of a character's moral and spiritual decline. Before imposing possession, the GM should:

\begin{enumerate}
\item Ensure the player understands and consents to the possibility
\item Provide multiple opportunities for the character to resist or break free
\item Make the descent into possession a gradual process with clear warning signs
\item Offer meaningful ways for other characters to help their companion
\item Respect the player's agency even when the character is compromised
\end{enumerate}

The path to possession should be paved with choices—each one a small surrender, each one making the next compromise easier. A character who deliberately seeks dark power should understand that they are playing with forces that demand payment in ways they may not anticipate.

\section*{The Ritual of Exorcism}

When possession takes root, when the shadow has grown too deep, there remains one path to salvation: the Ritual of Exorcism. But know this—exorcism is not a simple spell cast and forgotten. It is warfare. It is negotiation. It is sometimes surrender.

\textbf{Purpose:} To forcibly remove a possessing entity from a possessed subject

\textbf{Primary Caster:} Spirit 4+, Faith 3+, Lore 2+

\textbf{Helpers:} 2+ participants with Faith, Spirit, or relevant cultural/religious skills

\textbf{The Exorcism Encounter:}

Exorcism functions as a special encounter with its own clock mechanics. The ritual creates a contested space where the caster's will battles the possessing entity for control of the subject's soul.

\textbf{Exorcism Clock:} 2 + Subject's Tier in segments (minimum 4 segments)

Each segment represents a phase of spiritual combat requiring:
\begin{enumerate}
\item \textbf{Channel Phase (Spirit + Faith):} Primary caster and helpers focus their spiritual energy
\item \textbf{Weave Phase (Presence + Command):} Direct confrontation with the possessing entity
\item \textbf{Response Phase:} The entity fights back—participants are Exhausted and cannot act for One Moment
\end{enumerate}

\textbf{Entity Response Actions:}
During each Response Phase, the possessing entity may attempt actions that, if not resisted, advance the Exorcism Clock:
\begin{itemize}
\item \textbf{Whispers of Doubt:} Tempt participants with promises or threats
\item \textbf{Physical Manifestation:} Cause harm to participants or the subject
\item \textbf{Spiritual Assault:} Drain spiritual energy from participants
\item \textbf{Reality Distortion:} Create illusions or false visions to confuse the ritual
\end{itemize}

Participants may resist these actions with appropriate rolls, but failure allows the clock to advance by one segment.

\textbf{Defensive Support:}
Non-participating allies may concentrate on defense, providing +1 die to resistance rolls for each defender. However, defenders cannot take other actions during the exorcism.

\textbf{Ritual Failure:}
If the Exorcism Clock fills completely before the ritual succeeds, the exorcism fails with severe consequences:
\begin{itemize}
\item The possessing entity grows stronger (+1 die to all future possession attempts)
\item The subject suffers additional corruption (advance Sanity loss or equivalent)
\item All participants suffer Fatigue 2 from spiritual exhaustion
\item The entity gains knowledge of the participants' weaknesses
\end{itemize}

\textbf{For Player Characters:}
When a PC becomes possessed and exorcism fails, they gain complications based on the possessing entity's Tier:
\begin{itemize}
\item \textbf{Tier I-II Entities:} 1 CP banked each session from the entity's influence
\item \textbf{Tier III-IV Entities:} 2 CP banked each session from the entity's influence  
\item \textbf{Tier V+ Entities:} 3+ CP banked each session from the entity's influence
\end{itemize}

These banked CP represent the entity's ongoing influence and may manifest as:
\begin{itemize}
\item Compulsive behaviors or speech patterns
\item Nightmares or visions that provide the entity information
\item Uncontrolled moments where the entity acts through the character
\item Social consequences as others notice the character's changed nature
\end{itemize}

\shadow{Rukhir shivers, and I realize his feathers have gone dull. "Even I," he says quietly, "have felt the touch of things that would wear me like a glove. The difference, Veyla, is that I chose to be worn."}

\textbf{Player Agency Preservation:}
While possessed characters suffer ongoing complications, their players must always retain meaningful agency:
\begin{itemize}
\item Players choose when to resist the entity's influence
\item Major character decisions remain with the player
\item The entity cannot force the character to betray core relationships without resistance
\item Redemption paths should always remain possible, however difficult
\end{itemize}

The Ritual of Exorcism is not merely about casting spells—it is about the fundamental struggle between light and shadow, freedom and control, self and other. Approach it with the gravity it deserves, and remember that sometimes, the greatest mercy is to help a friend find peace, even if that peace requires them to let go of everything they were.

\textbf{Duration:} Variable encounter time, typically 20-30 minutes of game time per segment

\textbf{Casting Time:} Extended ritual requiring continuous focus and participation

\section*{Mist-Road Veil}
\textit{Draws a low fog that dulls footfalls and blurs silhouettes along a chosen path}

\textbf{Purpose:} Create concealing mist that aids stealthy movement along a specific route

\textbf{Primary Caster:} Wits 3+, Arcana 2+, Survival 2+

\textbf{Helpers:} 1-2 participants with Survival or Stealth skills

\textbf{Channel Phase (Wits + Survival):}
\begin{enumerate}
\item Gather water sources and cooling materials (damp cloth, morning dew, etc.)
\item Trace the path where mist is desired with light touch
\item Roll Wits + Survival for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Arcana):}
\begin{enumerate}
\item Channel moisture into the air along the designated path
\item Whisper words of concealment in the local tongue
\item Roll Wits + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} Mist is patchy or reveals more than it conceals; -1 die to Stealth rolls
\item \textbf{3-4 CP:} Mist attracts unwanted attention from creatures that hunt in fog; 2 CP added to next encounter
\item \textbf{5+ CP:} Mist becomes supernaturally thick, disorienting even the caster; party suffers -2 dice to navigation for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Creates concealing fog along a path up to 100 yards long for Significant Time
\item Party members moving along the path gain +2 dice to Stealth rolls
\item Sound is muffled by 50\% within the mist
\item Non-magical tracking becomes nearly impossible within the affected area
\end{itemize}

\shadow{Rukhir circles the area warily. "Mist hides the hunter," he caws, "but it also hides the hunted. Choose your prey wisely."}

\textbf{Duration:} Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (10-15 minutes of preparation and ritual)

\section*{Bell of Truce}
\textit{Rings a hospitality note that compels all within earshot to parley before violence}

\textbf{Purpose:} Create a magical compulsion for peaceful negotiation before combat

\textbf{Primary Caster:} Presence 4+, Spirit 3+, Command 2+

\textbf{Helpers:} 1-2 participants with Sway or Performance skills

\textbf{Channel Phase (Spirit + Command):}
\begin{enumerate}
\item Obtain a bell or chime of significant quality
\item Focus on the concept of hospitality and peaceful resolution
\item Roll Spirit + Command for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Performance):}
\begin{enumerate}
\item Ring the bell with specific intent and rhythm
\item Speak words of truce in the old tongue
\item Roll Presence + Performance for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The compulsion is weak or easily resisted; -1 die to Sway rolls for next social encounter
\item \textbf{3-4 CP:} The bell's influence extends beyond intended area, affecting neutral parties; 2 CP added to crowd-related complications
\item \textbf{5+ CP:} The compulsion backfires, making the caster appear weak or cowardly; -2 dice to Command rolls for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item All intelligent beings within 200 yards must make Spirit + Resolve (DV 3) to initiate violence for Some Time
\item Successful resistance grants +1 die to subsequent social rolls with that individual
\item Failure to parley first creates 2 CP backlash when violence finally erupts
\item The bell's tone marks a sacred space of temporary sanctuary
\end{itemize}

\shadow{Rukhir preens. "Even warriors must pause to hear beautiful sounds," he observes. "It is in that pause that wisdom often speaks."}

\textbf{Duration:} Some Time (approximately 30 minutes)

\textbf{Casting Time:} Some Time (ritual ringing and proclamation)

\section*{Lantern Weir}
\textit{Stitches light into a soft fence that guides crowds and resists panic surges}

\textbf{Purpose:} Create a barrier of light that guides movement and prevents chaotic rushes

\textbf{Primary Caster:} Wits 3+, Arcana 3+, Performance 2+

\textbf{Helpers:} 2-3 participants with Performance or Command skills

\textbf{Channel Phase (Wits + Performance):}
\begin{enumerate}
\item Gather multiple light sources (lanterns, candles, etc.)
\item Arrange participants in a line or curve to form the barrier
\item Roll Wits + Performance for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Weave the light sources together into a unified barrier
\item Speak words of guidance and protection
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The light barrier flickers or has gaps; -1 die to Crowd management rolls
\item \textbf{3-4 CP:} The barrier attracts attention from the wrong crowd; 2 CP added to next social encounter
\item \textbf{5+ CP:} The light becomes blinding rather than guiding; all within 50 yards suffer -1 die to perception rolls
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Creates a 50-foot barrier of soft, guiding light for Significant Time
\item Crowd movement is directed and calmed; +2 dice to Crowd management rolls
\item Panic effects are reduced by 2 CP for those within the barrier
\item The barrier provides dim illumination equivalent to moonlight
\end{itemize}

\shadow{Rukhir examines the light sources critically. "Light is a gentle master," he says, "but it demands respect from both those who wield it and those who would flee from it."}

\textbf{Duration:} Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (15-20 minutes of arrangement and ritual)

\section*{Oath of Shared Breath}
\textit{Binds the party's rhythm; fatigue spreads thinly among all who consent}

\textbf{Purpose:} Distribute fatigue and harm among willing party members

\textbf{Primary Caster:} Presence 4+, Spirit 3+, Bonds 2+

\textbf{Helpers:} All willing party members (2+ participants required)

\textbf{Channel Phase (Spirit + Bonds):}
\begin{enumerate}
\item All participants join hands or touch in a circle
\item Each participant speaks their willingness to share the burden
\item Roll Spirit + Bonds for primary caster, Spirit + Resolve for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Weave the participants' life forces together through ritual words
\item Create a shared rhythm of breath and heartbeat
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 CP:} The sharing is uneven, favoring some over others; one participant suffers +1 Fatigue
\item \textbf{3-4 CP:} The bond creates unwanted emotional connections; party suffers -1 die to social rolls with outsiders for Significant Time
\item \textbf{5+ CP:} The sharing becomes too intimate; participants can sense each other's thoughts and emotions for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item All willing participants share Fatigue and Harm equally for one session
\item Total Fatigue is divided by number of participants (rounded up)
\item Harm from attacks is distributed among all participants
\item Participants can sense each other's general health and location within 100 yards
\end{itemize}

\shadow{Rukhir tilts his head. "The heart that beats alone," he says, "can stop alone. But hearts that beat together... those are harder to silence."}

\textbf{Duration:} One full session

\textbf{Casting Time:} Some Time (ritual circle and shared oath)

	\section*{River's Favor}
	\textit{Coaxes currents to slacken at a ford or quay so boats and rafts pass cleanly}

	\textbf{Purpose:} Calm water currents to aid safe passage for watercraft

	\textbf{Primary Caster:} Wits 3+, Arcana 2+, Survival 3+

	\textbf{Helpers:} 1-2 participants with Survival or Pilot skills

	\textbf{Channel Phase (Wits + Survival):}
	\begin{enumerate}
	\item Touch the water at the point of desired calm
	\item Speak to the river's spirit or natural forces
	\item Roll Wits + Survival for primary caster, relevant skills for helpers
	\end{enumerate}

	\textbf{Weave Phase (Spirit + Arcana):}
	\begin{enumerate}
	\item Channel calming energy into the water
	\item Create a path of gentle currents
	\item Roll Spirit + Arcana for primary caster
	\end{enumerate}

	\textbf{Backlash:}
	\begin{itemize}
	\item \textbf{1-2 CP:} The calm is temporary or localized; -1 die to Pilot rolls
	\item \textbf{3-4 CP:} The river's natural flow is disrupted elsewhere; 2 CP added to next water-related encounter
	\item \textbf{5+ CP:} The ritual angers local water spirits; introduce a water-based supernatural threat
	\end{itemize}

	\textbf{Effect:}
	\begin{itemize}
	\item Creates a 100-foot stretch of calm water for Significant Time
	\item Watercraft gain +2 dice to Pilot rolls within the affected area
	\item Currents are reduced by 50\% within the calm zone
	\item Small debris and obstacles are pushed to the edges of the calm area
	\end{itemize}

	\shadow{Rukhir drinks from the river cautiously. "Water remembers promises," he says, "but it also remembers betrayals. Be gentle in your asking."}

	\textbf{Duration:} Significant Time (approximately 1 hour)

	\textbf{Casting Time:} Some Time (communion with the water and ritual work)

	\chapter{Rites of the Binding Arts}
	\label{chap:outer-rites}

	\lettrine[lines=3]{O}{}ut there, beyond the lamplight and the last safe mile of road, the world unfolds into angles that do not ask permission to exist. Some call them the outer spheres; I call them the \emph{places that look back}. If you would traffic with such places and the minds that dwell there, you will not do so by reciting tidy cantrips. You will bargain. You will invite. You will bind yourself as surely as you bind them.

	\shadow{Rukhir clicks his beak and tilts one midnight eye at you. ``She thinks she is ready to be \emph{kept},'' he croaks, ``but does she know what it means to be \emph{kept}?''}

	These pages gather what the temple fires failed to burn: names that answer when called, diagrams that hum under the skin, rites that bear the weight of attention. Do not mistake them for common spellwork. A rite is a door you open with your own breath and blood, a vow you sign with the hours of your life.

	\section*{A Single Thread to Hold}
	You may walk with a companion from beyond, and you may keep a book of forbidden knowledge. \textbf{But you may only align with one Patron at a time.} The mind that tries to serve two masters will find itself torn and many patrons are quite jealous.

	\section*{The Price the Powers Prefer}
	Every calling across the veil is answered with a ledger. You may not hear the scratching of the quill at first, but the ink dries all the same. Each time you Invoke a rite, the one you call \emph{notices}. Their notice is not love and not hate; it is gravity.

	\begin{verse}
	Power borrowed is power tallied; \\
	A boon spent is a bell rung. \\
	When the bell is full of echoes, \\
	Something comes to listen.
	\end{verse}

	When your tallies swell, the world arranges a lesson. Perhaps the sea decides it is done being polite. Perhaps a truth hounds you down the alley you thought was yours. Do not panic; do not run. Kneel, count the echoes, pay what must be paid, and move with care.

	\whisper{Do not hide your costs from yourself. Hide them from your enemies, perhaps, but never from your own ledger.}

	\section*{On Patrons and Their Proper Address}
	Call them Witness or Geometry, Sea or Serpent, Fire or Law, Bone or Dream, the Elsewhere or the Red Court—names are handles, not cages. Treat each with the grammar they prefer. Order listens to proofs. Shadows favor vows made without torches. Seas respect the traveler who brings a tide of their own.

	\shadow{``Say a thing as if the world already agrees,'' Rukhir advises, ``and half the time the world is too proud to argue.''}

	Each rite you will meet below is written with \emph{The Weaving} (how to call), \emph{The Price} (what will be marked), and \emph{The Whisper} (what answer tends to come first). Read all three. Then read them again in a different mood.

	\section*{The Mercy of Severance}
There are rites to bind, and there is one to \emph{unbind}. I place it at the end of this gathering, under a black thread and a warning bell. We call it \emph{Cut the Knot of Patronage}. It is not a betrayal to sever a cord that is choking you. It is a kind of piety.

\shadow{``Mercy is a sharp thing,'' says Rukhir, preening. ``Only the kind learn to use it.''}

Do not attempt severance in mockery or haste. Bring witnesses who will speak your name without flinching. Bring an emblem of how you first said \emph{yes}. Bring enough quiet to hear the cord go slack.

\section*{How to Use What Follows}
Begin with one thread only. Learn how it pulls when you pull it. When you Invoke, mark; when you are marked, answer. If you carry a Codex, \emph{tend it}. If you walk with a Thiasos, \emph{guard it}. When attention gathers like storm-smell on stone, step under an eave and wait for the first fat drop. Then choose whether to dance, or to shut the door and count your candles.

\whisper{The spheres are not cruel. They are vast. Vastness feels like cruelty to those who will not look up. Look up. Then look back down and mind your footing.}

If you are still reading, you are ready enough to begin. Turn the page, seeker. The doors do not open themselves—but they do not stay closed for those who learn the grammar of keys.

\section{Design Premise}
Runekeepers leverage \textbf{Rites}---structured, named bargains with alien patrons. A Rite is not a spell; it does not use Channel/Weave. Instead, it runs on \textbf{Boons} and a \textbf{Patron Obligation Clock} that measures attention and debt.

\subsection*{Access Ladder for Rites}
\begin{description}
  \item[Low-Power Rites:] \textbf{Requires:} Either a Thiasos Bond \emph{or} a relevant Codex. \textbf{Invoke:} 1 Boon. \textbf{Obligation:} 4-segment clock.
  \item[Standard Rites:] \textbf{Requires:} Both a Thiasos Bond \emph{and} a relevant Codex. \textbf{Invoke:} 1 Boon. \textbf{Obligation:} 5–6 segments.
  \item[High-Power Rites:] \textbf{Requires:} Thiasos Bond, relevant Codex, plus \textbf{Tier III} standing with your Patron. \textbf{Invoke:} \textbf{2 Boons}. \textbf{Obligation:} 7–8 segments.
\end{description}

\paragraph{Single Patron Bond}
A character can only be bound to \textbf{one patron at a time}. Your Thiasos and Codex are intrinsically tied to this bond. To switch patrons, a character must undergo the \textbf{Rite of Severance} (\S\ref{subsec:severance}), a costly and dangerous process that severs the old bond before a new one can be formed.

\paragraph{Shared Limits}
A Rite's clock may gain at most \textbf{Tier +2 segments} per scene (Invoke + Push It + Backlash). If a single Rite marks \textbf{2+} segments in a scene, your connection strains visibly—your Thiasos shows distress or your Codex grows hot to the touch.

\paragraph{Stacking}
Bonuses from \emph{scene-long} Rites do not stack with other scene-long bonuses from the same Patron. Apply the single best effect.

\section{Core Procedure: Invoking a Rite}
\label{sec:rite-procedure}
\textbf{Invoke:} Spend the required \textbf{Boon(s)} and mark \textbf{+1 segment} on that Rite's Obligation Clock.

While the Rite is in effect this scene, you may mark at most \textbf{two more} segments from the following triggers (each may occur once/scene per Rite):
\begin{enumerate}
  \item \textbf{Push It:} You use the Rite's ``\emph{Push It}'' rider (the bolder, alternative clause). \emph{+1 segment}.
  \item \textbf{Backlash:} A roll \emph{aided by the Rite} results in a \emph{Miss} or \emph{Success \& Cost}. \emph{+1 segment}.
\end{enumerate}
\textbf{Cap:} A single Rite can add at most \textbf{Tier +2 segments} in a scene.

\textbf{Clearing Segments:}
\begin{itemize}
  \item During play, you may spend \textbf{1 Boon} to clear \textbf{1 segment} on a Rite.
  \item During Downtime, a \emph{Maintained} \textbf{Codex} clears \textbf{1 segment} from one associated Rite.
  \item Tending to your \textbf{Thiasos} during Downtime (a specific activity) can clear \textbf{1 segment} from one Rite.
  \item \textbf{Ritual of Appeasement:} Spend \textbf{1 Boon} and dedicate \textbf{Significant Time} in Downtime to clear \textbf{1 segment} from a Patron's Obligation Clock (once per Patron per session, requires 2+ segments).
\end{itemize}

\textbf{Attention:} When a Rite's Obligation Clock fills, the Patron's attention manifests as a narrative complication. Start a new clock for the "Patron's Lesson."

\section{Character Options}
\subsection{Talent: Thiasos Bond (Familiar)}
\label{subsec:thiasos-bond}
\textbf{Cost:} 6 XP \hfill \textbf{Type:} General Talent \\
\textbf{Effect:} You bind a \emph{Thiasos}---a supernatural familiar keyed to your Patron. It may Assist actions involving that Patron's Rites. The Thiasos follows Familiar rules for Exposure/Harm.\\
\textbf{Upkeep:} Requires occasional tending (roleplayed interaction) to remain receptive. A neglected Thiasos may refuse to Assist.\\
\textbf{Risk:} If a Rite marks 2+ segments in a scene, the Thiasos is endangered: mark \emph{Exposure +1} or suffer \emph{Harm 1} on it.

\subsection{Asset: Codex of the Binding Art}
\label{subsec:codex-asset}
\textbf{Type:} Asset (Minor 4 XP / Standard 8 XP) \\
\textbf{Scope:} A Codex is dedicated to a specific Patron. A Minor Codex covers 1-2 Rites; a Standard Codex covers 3-4 related Rites.\\
\textbf{Upkeep:} The Codex uses the Asset Condition track:
\begin{quote}
\textbf{Maintained} (full), \textbf{Neglected} (−1 die to Invokes), \textbf{Compromised} (unavailable).
\end{quote}
In Downtime, you may \emph{Study the Codex} to restore it to \emph{Maintained} and clear \textbf{1 segment} on an associated Rite.

\subsection{Rite Costs (Learning)}
\textbf{Learn a Rite:} 4–12 XP. Low-Power (4-5 XP), Standard (7-9 XP), High-Power (12-14 XP). Requires the appropriate access (Thiasos or Codex for Low; both for Standard; both + Tier III for High).

\section{Rites by Patron}
Each Patron's rites are organized into three tiers of power. All Invocations cost \textbf{1 Boon} unless noted. Push It effects now require a meaningful cost or limitation.

\subsection{The Witness (Observation \& Truth)}
\paragraph{Eyes in the Shadow (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Obsidian focus; a drop of blood.
\textbf{Effect:} Auto-succeed on one mundane Notice this scene \emph{or} ask one yes/no about immediate surroundings.
\textbf{Push It:} Also gain +1 die on all investigation rolls this scene, but mark Exposure +1.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Unblinking Gaze (Low, 5 XP)} \emph{Duration: Hour; Range: Self; Stacking: No.}
\textbf{Materials:} A drop of ink placed on the eyelid.
\textbf{Effect:} Perfectly recall anything you consciously focus on within Near range for the duration. This grants +2 effect to a single Recall action or allows automatic success on memorizing a single page, map, or short sequence.
\textbf{Push It:} The memory becomes eidetic and can be perfectly reproduced, but you cannot form new memories for the next hour and suffer Fatigue 1.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Veil of the Unseen Observer (Standard, 7 XP)} \emph{Scene; Self; No.}
\textbf{Effect:} +2 dice to perception/investigation/recall; ask one specific factual question the GM must answer if knowable.
\textbf{Push It:} Pierce one mundane disguise/obfuscation once, but generate 1 CP (Spades) from the revelation.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Frozen Moment (Standard, 8 XP)} \emph{Duration: Instant; Range: Self; Stacking: No.}
\textbf{Materials:} A shard of glass or ice.
\textbf{Effect:} As a reaction to taking harm from a single, small object (e.g., an arrow, a thrown knife), you freeze it in motion for one beat, causing the attack to miss. The object hangs in stasis.
\textbf{Push It:} The frozen object can be gently redirected, but you suffer Fatigue 1 from the temporal strain.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Gaze of Absolute Truth (High, 10 XP)} \emph{Scene; Near; No.}
\textbf{Effect:} You cannot be lied to; see through illusions/disguises in Near; ask one ``true nature'' question about a magically altered subject.
\textbf{Push It:} Extend to allies in Close for one beat, but you become vulnerable to all forms of deception for the scene.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 6 segments.

\paragraph{Rite of the Perfect Record (High, 11 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A silvered lens or a pool of still water.
\textbf{Effect:} For the scene, you can perfectly replay the last 6 seconds of events within the zone as a silent, ghostly image. This can reveal hidden actions, confirm alibis, or expose lies. Alternatively, you may ask the GM one question about an event that occurred in the zone within the last hour.
\textbf{Push It:} The replay includes sound, but you must witness the replay yourself and cannot act during it.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 6 segments.

\subsection{The Sacred Geometry (Order \& Pattern)}
\paragraph{Find the Pattern (Low, 5 XP)} \emph{Action; Self; Yes (investigation only).}
\textbf{Materials:} A compass and a straightedge.
\textbf{Effect:} +1 die to decode patterns/codes/systems; re-roll one 1 on math/logic rolls.
\textbf{Push It:} Upgrade effect by one step on a single roll, but you become obsessed with finding patterns in unrelated phenomena for the scene and suffer -1 die to social rolls.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Ordered Step (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Walking in a perfect square.
\textbf{Effect:} Your movements become unnaturally precise. Ignore penalties from difficult terrain for walking. +1 die to any action requiring perfect physical calibration or balance.
\textbf{Push It:} You can walk a short distance across a fragile surface without causing a sound or breakage, but you must follow a geometrically perfect path and cannot deviate for the scene.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Thread the Loom of Chance (Standard, 7 XP)} \emph{Action; Self; No.}
\textbf{Materials:} A set of weighted dice or a balanced scale.
\textbf{Effect:} Re-roll up to \emph{two dice}
\textbf{Push It:} Treat one zone tag as favorable for this action, but you must accept an equal but opposite consequence later in the scene.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Golden Ratio (Standard, 7 XP)} \emph{Duration: Scene; Range: Touch; Stacking: No.}
\textbf{Materials:} A measured length of string, divided according to the golden ratio.
\textbf{Effect:} Touch one object no larger than a door. For the scene, its structure is optimized. A door becomes harder to break (+1 effect to resist), a weapon strikes truer (+1 die to its next attack), or a tool provides +1 effect on its next use.
\textbf{Push It:} The optimization affects two similar, connected objects, but the effect on the second object is half strength.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rewrite the Fundamental Equation (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A complex geometric diagram drawn at the zone's center.
\textbf{Effect:} Declare one physics/magic rule different within a zone (no instant kills; GM veto for scene-enders). Once/scene, downgrade one \emph{Miss} to \emph{Success \& Cost}.
\textbf{Push It:} Affect a second adjacent zone for one beat, but the change creates an unstable paradox that generates 2 CP.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Platonic Form (High, 13 XP)} \emph{Duration: Scene; Range: Near; Stacking: No.}
\textbf{Materials:} A perfectly shaped crystal or gem.
\textbf{Effect:} You perceive the ideal form behind reality. Choose one: Gain +2 dice to see through illusions and disguises; or, when crafting or repairing, you achieve a flawless result, granting the object +1 effect on its first use.
\textbf{Push It:} You can impose this perfect form for a moment, causing a flawed object to function perfectly once, but the object shatters afterward.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Ikasha, She Who Sleeps (Latent Potential \& Shadow)}
\paragraph{Touch the Umbral Veil (Low, 4 XP)} \emph{Action; Self; Yes (stealth only).}
\textbf{Materials:} A piece of black cloth.
\textbf{Effect:} Start \emph{Controlled} on one Stealth roll or add +1 effect to hiding/moving quietly.
\textbf{Push It:} Gain brief shadow-muffling (ignore one noisy tell), but you leave behind a shadow-double that mimics your movements.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Whispering Shade (Low, 5 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} Extinguishing a candle.
\textbf{Effect:} Cause shadows in the zone to subtly move and distort. This can be used to grant +1 die to a Create Diversion action or to impose a −1 die penalty on a single enemy's concentration-based action.
\textbf{Push It:} The shadows form a brief, terrifying shape, but the shadows remember your face and may seek you out later.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Draw from the Umbral Reservoir (Standard, 8 XP)} \emph{Action; Self or Ally; No.}
\textbf{Materials:} A vial of water from a moonless night.
\textbf{Effect:} +2 dice to stealth/deception/inner-reserve \emph{or} clear \emph{Fatigue 1} from target.
\textbf{Push It:} Also gain one free escape attempt, but you must pay the favor forward by helping another escape within the next scene.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Secret Keeper's Burden (Standard, 9 XP)} \emph{Duration: Instant; Range: Touch; Stacking: No.}
\textbf{Materials:} A lock of the target's hair or a personal item.
\textbf{Effect:} Compel a touched creature to truthfully answer one specific, direct question. They may be cryptic, but cannot lie. The GM may allow a Resolve roll to resist if the question touches on a deeply protected secret.
\textbf{Push It:} You learn not only the answer, but a key emotional context, but the target learns one of your secrets in return.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Become the Shadow Itself (High, 12 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Standing in a pool of absolute darkness.
\textbf{Effect:} Intangible to mundane harm; pass through small openings; +2 dice to Stealth; auto-succeed one escape. Cannot manipulate normal objects.
\textbf{Push It:} Interact with a single bound object once, but you become partially corporeal and vulnerable for one beat.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Silent Council (High, 13 XP)} \emph{Duration: Scene; Range: Near; Stacking: No.}
\textbf{Materials:} Whispering a question into a darkened corner.
\textbf{Effect:} The shadows themselves become your informants. You may ask up to three yes/no questions about events that have occurred within the area in the last 24 hours, receiving answers from the collective memory of the darkness.
\textbf{Push It:} You may ask one of the questions about an event that is currently happening out of your direct sight, but the shadows demand a favor in return to be paid within the week.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Khemesh, the Kraken (Crushing Depths)}
\paragraph{Grasp of the Minor Depth (Low, 5 XP)} \emph{Action; Near; Yes (move/drag only).}
\textbf{Materials:} A handful of wet sand.
\textbf{Effect:} +1 effect to pull/drag/pressure \emph{or} on a hit against you, force attacker to generate 1 CP (their suit).
\textbf{Push It:} Impose \emph{Off-Balance} (minor condition) on a target you move, but you become briefly waterlogged and suffer −1 die to your next action.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Briny Taste (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A sip of salt water.
\textbf{Effect:} You exert a subtle, constant pressure. Foes in Close range feel a palpable sense of weight, suffering −1 die on initiative or quick-draw actions. You gain +1 die to resist being moved or grappled.
\textbf{Push It:} A single foe in Close range is visibly slowed, but the pressure affects you as well, causing Fatigue 1.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Command the Grasp of the Abyss (Standard, 8 XP)} \emph{Action; Near; No.}
\textbf{Materials:} A deep-sea shell held to the ear.
\textbf{Effect:} Force one foe into \emph{Desperate} position \emph{or} +2 effect to destruction.
\textbf{Push It:} Spread \emph{Desperate} to a second engaged foe, but you must make a Spirit + Resolve test (DV 2) or become momentarily overwhelmed by the depths.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Inexorable Tide (Standard, 7 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} Tracing a wave pattern on the ground.
\textbf{Effect:} The environment becomes sluggish and heavy. All movement within the zone costs 1 extra AP, or the first movement each character makes is treated as one position step worse (Controlled→Risky, Risky→Desperate).
\textbf{Push It:} The effect is particularly strong against one named target, affecting them twice, but the tide turns against you if you remain in the zone.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Command the Abyssal Maw (High, 12 XP)} \emph{Action; Zone; No.}
\textbf{Materials:} The sound of a great wave crashing.
\textbf{Effect:} Force all opponents in scene into \emph{Desperate} if fictionally justified \emph{or} +3 effect to pure destruction.
\textbf{Push It:} Collapse cover or footing in one zone, but you become partially submerged in the conceptual depths and suffer −1 die to social actions for the scene.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Kraken's Embrace (High, 14 XP)} \emph{Duration: Scene; Range: Near; Stacking: No.}
\textbf{Materials:} The bone of a large sea creature.
\textbf{Effect:} You create a zone of crushing deep-sea pressure. All within it suffer −2 dice to physical actions and cannot take two actions in a beat. Those who start their turn in the zone must resist or gain the \emph{Stunned} condition.
\textbf{Push It:} The pressure becomes tangible, allowing you to make a crushing attack against everyone in the zone, but you must concentrate fully and cannot take other actions.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Raéyn of the Sea (Tides \& Flux)}
\paragraph{Read the Gentle Current (Low, 4 XP)} \emph{Action; Self; Yes (nav only).}
\textbf{Materials:} A floating leaf or a weather vane.
\textbf{Effect:} +1 die to navigation/weather/flow \emph{or} shift 1 CP (minor) to a different suit.
\textbf{Push It:} Grant an ally +1 die on the next travel check, but you become temporarily disoriented and suffer −1 die to your next navigation roll.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Shifting Sand (Low, 5 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A handful of sand or a changing cloud.
\textbf{Effect:} You adapt to changing circumstances. Once per scene, when a situation changes unexpectedly (a new foe arrives, the weather turns), you may take a moment to reassess, granting yourself +1 die on your next action related to the new situation.
\textbf{Push It:} An ally in Close range also gains this benefit, but you must guide them through the adaptation, limiting your own actions.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Ride the Tides of Fate (Standard, 7 XP)} \emph{Action; Self; No.}
\textbf{Materials:} A coin flipped three times.
\textbf{Effect:} Shift one declared CP spend to a different suit \emph{or} +1 die to water travel/change-navigation.
\textbf{Push It:} Reduce one foe's effect by one step if they contest movement, but the tides remember and may turn against you later.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Ebbing Vigor (Standard, 8 XP)} \emph{Duration: Scene; Range: Near; Stacking: No.}
\textbf{Materials:} A retreating wave or a fading echo.
\textbf{Effect:} You can weaken a target's resolve. A single character in Near range suffers −1 die to Resolve rolls and their emotions become more volatile (GM may shift generated CP towards Hearts or Spades).
\textbf{Push It:} The target also suffers −1 die on their next action based on their current dominant emotion, but their emotional state becomes unpredictable and may affect allies as well.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Call the Storm's Judgment (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Shouting into the wind.
\textbf{Effect:} Name a target; environment opposes them: −1 die to actions and they generate CP on success involving exertion. \emph{Or} clear all \emph{Fatigue} in Near allies.
\textbf{Push It:} Extend penalty to a second, related target, but the storm's attention is drawn and may affect unintended targets.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Uncharted Course (High, 13 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A map with a blank spot.
\textbf{Effect:} You become a nexus of beneficial chance. Once per beat, you may reroll a single die for any roll you make or an ally in Close range makes. Alternatively, you may force an enemy to reroll a single die on their successful action.
\textbf{Push It:} You may use this ability twice in a single beat, but you must choose which rolls to influence before seeing the results.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{The Clockwork Demon (Forbidden Tech \& Entropy)}
\paragraph{Whisper to the Gears (Low, 4 XP)} \emph{Action; Near; Yes (simple mech).}
\textbf{Materials:} A drop of oil or a loose screw.
\textbf{Effect:} +1 die to simple mechanisms \emph{or} make a broken simple device work once.
\textbf{Push It:} Also gain +1 effect on this fix, but the device becomes temperamental and may fail at a critical moment.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).} \textbf{Note:} Consider upgrading to +2 effect for more impact.
\paragraph{Rite of the Unseen Flaw (Low, 5 XP)} \emph{Duration: Instant; Range: Near; Stacking: No.}
\textbf{Materials:} Tapping a complex device three times.
\textbf{Effect:} You intuitively understand a point of failure in a single mechanical or social system you can see. On your next action to exploit that flaw, you gain +2 effect.
\textbf{Push It:} The flaw manifests immediately as a minor, distracting failure, but you can also exploit a second, related flaw with +1 effect.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Wind the Gears of the Proscribed (Standard, 9 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A forbidden tool or diagram.
\textbf{Effect:} Temporarily repair/enhance a device for the scene \emph{or} +2 dice to Crafts/Engineering with forbidden methods.
\textbf{Push It:} Add one extraordinary flourish, but the device becomes unstable and may explode or malfunction violently.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Calculated Breakdown (Standard, 8 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} Scattering a handful of rust or sand.
\textbf{Effect:} Accelerate entropy in a zone. Complex devices are more likely to jam or fail (−1 die to operate), and structures creak and strain. You gain +1 effect to any action that aims to break or sabotage something in the zone.
\textbf{Push It:} A single device of your choice in the zone suffers a catastrophic failure, becoming unusable, but you can scavenge useful components from the wreckage.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Wind the Engine of Forbidden Knowledge (High, 15 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A core of unstable energy.
\textbf{Effect:} Dramatically repair/create a complex device for a scene \emph{or} +3 dice to forbidden engineering.
\textbf{Push It:} Device exceeds specs for one action, but the unstable energy creates a hazardous field that affects all nearby.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Final Scrap (High, 14 XP)} \emph{Duration: Instant; Range: Touch; Stacking: No.}
\textbf{Materials:} A hammer blow or a final command.
\textbf{Effect:} You touch a complex object or structure (up to the size of a small building) and command it to fail. It suffers an immediate, catastrophic, and irreversible breakdown. This is not an explosion, but a rapid dissolution into its component, useless parts.
\textbf{Push It:} The breakdown is contained and quiet, but the released energy must go somewhere and affects the surrounding area.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Isoka, Angel of Serpents (Change \& Shedding)}
\paragraph{Loosen the Old Skin (Low, 4 XP)} \emph{Scene; Self; Yes (resist only).}
\textbf{Materials:} A discarded snakeskin or a loose thread.
\textbf{Effect:} +1 die to resist an ongoing Condition this scene \emph{or} re-roll one 1 on escape/evasion.
\textbf{Push It:} Also ignore one minor movement penalty, but you leave behind a piece of your old self that others can use against you.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Subtle Shift (Low, 5 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Shifting a small object from one pocket to another.
\textbf{Effect:} Your demeanor becomes fluid. +1 die to Deceive rolls to appear as a slightly different social class or profession. Alternatively, +1 effect to blend into a new crowd or environment.
\textbf{Push It:} The change is convincing enough to bypass a single, minor security check, but you must maintain the deception and cannot easily return to your original identity.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Shed the Skin of the Former Self (Standard, 8 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A full change of clothes and a new mannerism.
\textbf{Effect:} +2 dice to resist a named ongoing Condition; once/session declare a minor physical contingency retroactively.
\textbf{Push It:} Clear a \emph{temporary} identity-based minor Condition, but your former identity becomes aware of the change and may seek you out.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Forked Tongue (Standard, 7 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A deliberate, harmless lie told to a mirror.
\textbf{Effect:} Your words become ambiguously persuasive. When you Sway or Command, you may choose to have a successful roll generate \emph{Diamonds} (leverage) instead of the usual CP, as your words create useful misunderstandings or plausible deniability.
\textbf{Push It:} A specific, carefully worded lie you tell this scene is accepted as truth, but the truth you've displaced seeks to return and complicate matters.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Complete Metamorphosis (High, 12 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A complete and total change of identity, including personal effects.
\textbf{Effect:} Full appearance/voice change; start \emph{Controlled} on Deceive/Stealth; once/scene declare minor contingency retroactively.
\textbf{Push It:} Also spoof scent/biometric for one check, but your original identity becomes partially unmoored and may act independently.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Cast-Off History (High, 13 XP)} \emph{Duration: Extended; Range: Self; Stacking: No.}
\textbf{Materials:} Burning all records of your old life.
\textbf{Effect:} You perform an extended ritual to shed a past identity. Upon completion, all mundane records and common knowledge of your former self become hazy and unreliable. Those seeking you via that identity suffer −2 dice to their efforts. This does not affect magical scrying or those who knew you intimately.
\textbf{Push It:} The ritual creates a false but plausible "death" for your old identity, but someone who knew you well may sense the deception.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Inaea, Angel of the Spider (Webs \& Fate)}
\paragraph{Tie a Simple Knot (Low, 4 XP)} \emph{Action; Near; Yes (link once).}
\textbf{Materials:} A single strand of thread.
\textbf{Effect:} Declare two minor events linked; force 1 CP (suit chosen by GM) on a foe \emph{or} grant +1 die later to a related roll.
\textbf{Push It:} Pre-load the +1 as a held advantage this scene, but the web of fate becomes more complex and may entangle unintended targets.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Tangled Thread (Low, 5 XP)} \emph{Duration: Scene; Range: Near; Stacking: No.}
\textbf{Materials:} Gently tugging on a web or net.
\textbf{Effect:} You create a minor, invisible snare of fate. Designate a small area (e.g., a doorway, a patch of floor). The first person to enter that area suffers a minor stumble or fumble, imposing −1 die on their next action.
\textbf{Push It:} The snare is strong enough to briefly entangle, allowing an ally a free attack, but the entanglement affects all who enter, friend or foe.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Weave the Strand of Inevitability (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Weaving three different colored threads together.
\textbf{Effect:} Link two actions: when A moves, B becomes vulnerable/entangled; typically forces CP or grants +2 dice to prediction/setup.
\textbf{Push It:} Name a delayed beat where the link snaps tight, but the strain on the web affects you as well.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Weaver's Glance (Standard, 7 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Watching a spider build its web.
\textbf{Effect:} You see the connections between people. For the scene, you can easily discern who holds influence over whom in a social situation, and you gain +1 die to rolls to exploit those connections (Sway, Command, or Blackmail).
\textbf{Push It:} You discern one specific, hidden obligation or debt, but the web of connections becomes more visible to others as well.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Weave the Tapestry of Inevitability (High, 13 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A complex knotwork pattern drawn on the ground.
\textbf{Effect:} Name two targets; successes by one impose consequences on the other for the scene.
\textbf{Push It:} Add a third minor linkage for one beat, but the complexity of the web makes it harder to control.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Grand Design (High, 14 XP)} \emph{Duration: Extended; Range: Zone; Stacking: No.}
\textbf{Materials:} A map of the location with threads pinned to key points.
\textbf{Effect:} Over an extended period, you weave a "script" for a social or investigative encounter within a location. Once set, you can once per scene declare that a character's action plays directly into your planned narrative, allowing you to control their position for their next action (e.g., making a cautious character \emph{Desperate} or a hostile one \emph{Controlled}).
\textbf{Push It:} The script includes a hidden "trap," forcing a character of your choice to generate \emph{Spades} CP on their next success, but the web of fate becomes more tangled and affects you as well.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Mor'iraath, Demon of Fire (Consumption)}
\paragraph{Spark of the Inner Flame (Low, 4 XP)} \emph{Action; Self; Yes (destruction/passion).}
\textbf{Materials:} A hot coal or a pinch of sulfur.
\textbf{Effect:} +1 effect to destruction/passion \emph{or} ignore \emph{Fatigue 1} from a brief exertion.
\textbf{Push It:} Add a searing tell (may shift one CP suit to Clubs), but you suffer minor burns.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Forge-Heart (Low, 5 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A hot coal held to the chest.
\textbf{Effect:} Ignore environmental penalties from cold; cause flammable objects you touch (paper, dry cloth) to smolder; add +1 effect to actions involving heating or melting.
\textbf{Push It:} Your body radiates enough heat to become a minor source of warmth and light, but you suffer Fatigue 1 from the internal heat.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Kindle the Ember of Annihilation (Standard, 7 XP)} \emph{Action; Near; No.}
\textbf{Materials:} A shouted oath or a broken token.
\textbf{Effect:} Next damaging action gains \emph{Severe} if it hits \emph{or} +2 effect to pure destruction.
\textbf{Push It:} Ignite a hazardous tag in the scene, but the fire may spread beyond your control.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of Consuming Zeal (Standard, 8 XP)} \emph{Duration: Instant; Range: Near; Stacking: No.}
\textbf{Materials:} A passionate, convincing argument.
\textbf{Effect:} When you make a direct threat or declaration of intent, the target must resist with Resolve or be shaken, suffering −1 die on their next social action against you and generating \emph{Spades} CP on a success.
\textbf{Push It:} The intimidation spreads, applying a lesser version to a second, closely associated target, but your own passion becomes consuming and you suffer −1 die to your next social action.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Unleash the Ember of Total Annihilation (High, 13 XP)} \emph{Action; Zone; No.}
\textbf{Materials:} A significant sacrifice thrown into a fire.
\textbf{Effect:} Next damaging action gains \emph{Critical} if it hits \emph{or} +3 effect to catastrophic destruction; may ignore first two \emph{Fatigue} from one burst.
\textbf{Push It:} Shockwave escalates collateral one step, but the fire consumes part of your own essence and you suffer Fatigue 1.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Phoenix's Pyre (High, 14 XP)} \emph{Duration: Scene; Range: Touch; Stacking: No.}
\textbf{Materials:} An object of great personal value.
\textbf{Effect:} You ritually destroy something meaningful (a weapon, a tool, a token) in a controlled conflagration. In return, you or an ally touched by the flames may clear all \emph{Fatigue} and one major physical Condition, emerging refreshed but marked by soot and the smell of smoke.
\textbf{Push It:} The renewal also grants +1 die to all actions for the next scene, but the fire's hunger grows and demands greater sacrifices in the future.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Mykkiel, Angel of Law (Edict \& Process)}
\paragraph{Invoke the Letter of the Law (Low, 4 XP)} \emph{Action; Self; Yes (authority only).}
\textbf{Materials:} A scroll or a badge of office.
\textbf{Effect:} +1 die when acting as authority/enforcing a known rule \emph{or} force a successful minor action against you to either accept a small narrative consequence or re-roll at −1 die.
\textbf{Push It:} Name the statute; gain +1 effect on compliance, but you become bound by the same strict interpretation.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Ordered Task (Low, 5 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A checklist or a set of instructions.
\textbf{Effect:} You impose perfect order on a single, multi-step mundane process (e.g., searching a room, compiling reports, crafting a simple item). The task is completed in half the normal time, or you gain +1 effect on the final outcome.
\textbf{Push It:} The efficiency is contagious, granting one ally involved in the task a similar benefit, but you become obsessed with order and suffer −1 die to creative or spontaneous actions.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Proclaim the Edict of the Justiciar (Standard, 8 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Speaking in a formal tone of command.
\textbf{Effect:} +1 die to Command/Sway while enforcing a known law \emph{or} force a foe who just succeeded against you to ``pay the price'' (minor immediate consequence or re-roll at −1 die).
\textbf{Push It:} Your edict covers allies in Close for one exchange, but you become the focus of attention and all social actions against you are treated as if you were the primary target.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Unbreakable Vow (Standard, 7 XP)} \emph{Duration: 24 Hours; Range: Near; Stacking: No.}
\textbf{Materials:} A handshake or a signed agreement.
\textbf{Effect:} You witness a promise made between two parties. If either party knowingly breaks the vow, they are afflicted with the \emph{Guilty} condition (−1 die to all actions until they atone) until they make amends.
\textbf{Push It:} The condition is more severe, also causing them to generate \emph{Spades} CP on successful actions until atonement, but you become partially responsible for ensuring the vow is kept.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Declare the Judgement of the Justiciar (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Striking a gavel or a shield.
\textbf{Effect:} +2 dice to authority actions; opponents who violate clear principles generate \emph{Hearts} CP; once/scene, force a foe to accept a significant immediate consequence or re-roll at −2 dice.
\textbf{Push It:} Sanctify the venue for one scene tag (lawful), but you become the arbiter of justice and must deal with the consequences of your judgments.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Final Decree (High, 13 XP)} \emph{Duration: Instant; Range: Near; Stacking: No.}
\textbf{Materials:} A sealed verdict or an unbreakable chain.
\textbf{Effect:} You pronounce a magical judgment on a single target who has clearly broken a major, established law. The target must resist with Resolve or be paralyzed by magical chains for one beat, and all allies gain +1 effect on their first attack or action against the target.
\textbf{Push It:} The judgment is absolute; the target cannot benefit from protective magic for the duration of the scene, but the weight of absolute justice affects you as well and you suffer −1 die to mercy-based actions for the scene.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Varnek Karn, the Bone King (Death \& Dominion)}
\paragraph{Whisper to Restless Spirits (Low, 4 XP)} \emph{Action; Near; Yes (death sites).}
\textbf{Materials:} A bone fragment or a pinch of grave dust.
\textbf{Effect:} +1 die to recent-death investigation \emph{or} ask one yes/no about a death in the last day.
\textbf{Push It:} Sense nearest unquiet grave, but the spirits take note of your interest and may seek you out later.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Grave's Chill (Low, 5 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Holding a cold stone to your heart.
\textbf{Effect:} You emanate an aura of death. Animals are uneasy around you, and you gain +1 die to Intimidate rolls. You can also cause nearby plants to wilt slightly.
\textbf{Push It:} A single, small creature must flee from you in panic, but the aura of death clings to you and affects your interactions with the living.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Command the Restless Dead (Standard, 9 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A skull or a funerary bell.
\textbf{Effect:} Command minor undead up to your \emph{Arcana} (treat as Cap 3 follower swarm; Spades-leaning complications) \emph{or} +2 dice to resist disease/poison/necromancy or step down one undead-themed Harm.
\textbf{Push It:} Add a simple tactic to the swarm for one beat, but the undead become more aggressive and harder to control.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Soul's Price (Standard, 8 XP)} \emph{Duration: Instant; Range: Touch; Stacking: No.}
\textbf{Materials:} A silver coin placed on a corpse's eyes.
\textbf{Effect:} You can ask a single, recently deceased corpse (dead less than an hour) one question about the moments leading to its death. The corpse answers with its dying thoughts, though the memory may be fragmented.
\textbf{Push It:} The corpse can also answer one question about a secret it knew in life, but the act disturbs other nearby spirits.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Awaken the Army of Bones (High, 13 XP)} \emph{Extended; Zone; No.}
\textbf{Materials:} A banner made of bones raised over a graveyard.
\textbf{Effect:} Command an undead force (Cap 5 equivalent) for an extended conflict; concentration required \emph{or} +3 dice to death magic and immunity to fear from undead this scene.
\textbf{Push It:} One formation maneuver at Great effect, but the army's hunger grows and it may turn on you if not fed regularly.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Eternal Guardian (High, 14 XP)} \emph{Duration: Permanent; Range: Touch; Stacking: No.}
\textbf{Materials:} The intact skeleton of a loyal creature or warrior.
\textbf{Effect:} You animate a single skeleton to serve as a permanent, mindless guardian for a location or object you designate. It has the stats of a Cap 2 follower and will obey simple commands related to its guard duty. It can be destroyed, but not turned.
\textbf{Push It:} The guardian is exceptionally sturdy, effectively Cap 3 for the purposes of resilience and combat, but it becomes partially sentient and may develop its own agenda over time.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Mab, She of the Thorned Crown (Twilight Bargains)}
\textit{And speak not Nine to Mab...}
\paragraph{Fae Fortune's Favor (Low, 4 XP)} \emph{Action; Self; Yes (luck/social).}
\textbf{Materials:} A four-leaf clover or a laugh.
\textbf{Effect:} Re-roll one die showing 1–2 on luck/charm \emph{or} start \emph{Controlled} on one whimsical charm/deceit.
\textbf{Push It:} Name a fae quirk that colors the scene, but you become temporarily bound by fae logic and cannot act directly against it.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Thorny Promise (Low, 5 XP)} \emph{Duration: 24 Hours; Range: Near; Stacking: No.}
\textbf{Materials:} A spoken pact between two willing parties, sealed with a pricked finger.
\textbf{Effect:} A simple, verbal agreement is made magically binding. If either party knowingly breaks the exact terms, they suffer a minor but persistent consequence (e.g., \emph{Itchy} condition, −1 die on a specific type of action).
\textbf{Push It:} The consequences of breaking the pact become more severe, but Mab takes note of the bargain and may demand a favor in return.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Bargain Beneath the Hawthorn (Standard, 7 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A token exchanged under a tree.
\textbf{Effect:} +2 dice to Persuasion/Deceive if phrased as a formal bargain; once/session retroactively declare a minor owed favor with an on-scene NPC (GM defines their take).
\textbf{Push It:} Gain an extra +1 die for a particularly clever twist, but the bargain creates an equal but opposite obligation that must be fulfilled.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Glamour's Veil (Standard, 8 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A small mirror or a lock of hair from the target.
\textbf{Effect:} Choose one individual you can see. You appear maximally appealing or intimidating to them. Gain +2 dice on a single Sway or Command roll against that target.
\textbf{Push It:} The effect is strong enough to briefly override a known fact or strong first impression, but the glamour affects your own perception as well.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Courtship with the Thorned Queen (High, 13 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A gift of exquisite beauty or terrible cost.
\textbf{Effect:} +2 dice to all social interactions; force one NPC who opposes you to generate \emph{Hearts} CP; once, declare an action ``blessed by the Court'' (spectacular success mirrored by a later opposite consequence).
\textbf{Push It:} Temporary immunity to social manipulation for one beat, but the Court's attention is drawn and they may demand a price for the favor.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Unseelie Audience (High, 14 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} A circle of mushrooms and a spoken invitation to the Court.
\textbf{Effect:} You temporarily overlay the scene with the influence of Mab's court. Social interactions are governed by strict, arcane rules of etiquette. You know these rules perfectly, granting you +2 effect to all social rolls. Those who break the rules suffer a minor supernatural consequence (a loss of voice, a twisted limb) chosen by the GM.
\textbf{Push It:} You can name one specific rule that an opponent is certain to break, triggering the consequence immediately, but you must also abide by all the rules and cannot act outside them.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Nidhoggr, the World-Worm (Dreaming Antiquity)}
\paragraph{Glimpse the Ancient's Shadow (Low, 5 XP)} \emph{Action; Near; Yes (lore sense).}
\textbf{Materials:} A fossil or a piece of weathered stone.
\textbf{Effect:} +1 die to ancient history \emph{or} sense ancient auras/sites in Near.
\textbf{Push It:} Briefly glean a single historic image, but the ancient knowledge disturbs your understanding of the present.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Dusty Tome (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Reading from a forgotten text.
\textbf{Effect:} The weight of ages settles on you. You gain +1 die to resist magical charms and illusions, as they seem fleeting compared to eternal truths. Conversely, you suffer −1 die to actions involving modern technology or fast-paced events.
\textbf{Push It:} You can identify the approximate age of any object you touch, but you become temporarily disconnected from contemporary concerns.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Drink from the Dreaming Deep (Standard, 9 XP)} \emph{Action; Self; No.}
\textbf{Materials:} Sleeping atop a historic site.
\textbf{Effect:} Learn one hidden fact (ask a specific question; suffer \emph{Fatigue 1} and mark \emph{Exposure +1}) \emph{or} +2 dice to forbidden/ancient lore or reality-warping.
\textbf{Push It:} Also gain an omen, but the ancient knowledge comes with a curse that affects your dreams.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Stone-Sleeper's Murmur (Standard, 8 XP)} \emph{Duration: Scene; Range: Touch; Stacking: No.}
\textbf{Materials:} Pressing your ear against a very old wall or stone.
\textbf{Effect:} The stone shares its memories. You can ask the GM one question per beat about events that have occurred in direct contact with that stone within the last century. The answers are vague, sensory impressions (e.g., "many cold footsteps," "the warmth of a great fire").
\textbf{Push It:} You receive a clear, brief image related to one of the answers, but the stone's memories become part of your own and may surface at inconvenient times.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Dive into the World-Worm's Dreaming (High, 14 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A deep trance induced by ancient incense.
\textbf{Effect:} Ask three factual questions about past/hidden truth; suffer \emph{Fatigue 2} and \emph{Exposure +2} \emph{or} +3 dice to deep reality work and one \emph{Miss→Success \& Cost} per scene.
\textbf{Push It:} Share one glimpse with an ally, but the ancient knowledge affects both of you and creates shared visions.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Awakened Chronicle (High, 15 XP)} \emph{Duration: Extended; Range: Zone; Stacking: No.}
\textbf{Materials:} Inscribing a forgotten history onto a suitable surface.
\textbf{Effect:} You perform a ritual to make a location "remember" a past event. Over an extended period, you can cause a ghostly replay of a historical scene to manifest visibly and audibly for all present. This is only a memory and cannot be interacted with, but it reveals the truth of what occurred.
\textbf{Push It:} The replay includes the emotional state of one key participant, revealed as a chilling aura or a warm glow, but the memory becomes partially real and affects the present.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{The Outsider (Elsewhere \& Intrusion)}
\paragraph{Peek Through the Veil (Low, 5 XP)} \emph{Action; Self; Yes (env penalty).}
\textbf{Materials:} A key that opens nothing.
\textbf{Effect:} Treat one environmental/position penalty as one step less severe \emph{or} +1 die to detect planar auras.
\textbf{Push It:} Reveal a nearby exit not previously obvious, but the exit leads to an unpredictable location.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Unseen Door (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Marking a doorway with an impossible angle.
\textbf{Effect:} You sense all points of entry and egress within Near range, even hidden ones. You gain +1 die to find hidden doors or to barricade existing ones.
\textbf{Push It:} You can cause one small, unlocked door to appear locked and stuck for the scene, but you also create an unintended exit that leads to an unknown location.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Step Sideways (Standard, 8 XP)} \emph{Action; Self; No.}
\textbf{Materials:} A reflection that moves on its own.
\textbf{Effect:} Treat the next zone you enter as having a favorable environmental aspect for one action \emph{or} +2 dice to extradimensional/teleportation/resisting planar effects.
\textbf{Push It:} Also slide one pursued ally with you, but the dimensional shift affects your perception and you suffer −1 die to your next action.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Borrowed Perspective (Standard, 7 XP)} \emph{Duration: Scene; Range: Near; Stacking: No.}
\textbf{Materials:} Looking at a location through a piece of stained glass.
\textbf{Effect:} You can project your senses to a nearby location you have visited before (within the same building or complex). You can see and hear from that spot as if you were there, but you are physically vulnerable and unaware of your surroundings.
\textbf{Push It:} You can whisper a short message that can be faintly heard at the target location, but your attention is split and you suffer −1 die to all actions in your current location.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Tear the Fabric of Reality (High, 14 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A shard of impossible geometry.
\textbf{Effect:} For one action, ignore all environmental/position penalties; create a temporary portal to a similar location within Near (GM sets exact destination) \emph{or} +3 dice to planar magic; immunity to non-magical environmental effects this scene.
\textbf{Push It:} Leave a short-lived echo (hazard tag), but the tear in reality attracts unwanted attention from otherworldly entities.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Nowhere House (High, 15 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} Defining a space that is neither inside nor outside.
\textbf{Effect:} You create a temporary, extradimensional pocket space large enough to hold a few people. It is safe from the outside world but has no exits other than the one you create. Time passes strangely inside (an hour outside is a minute inside, or vice versa, GM's choice).
\textbf{Push It:} The space can store a small, non-living object indefinitely, retrieving it later from any similar pocket space you create, but the space becomes semi-permanent and may develop its own rules over time.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{Maelstreus, the Red Prince (Infernal Deals)}
\paragraph{Whisper of the Red Court (Low, 3 XP)} \emph{Action; Self; Yes (temptation).}
\textbf{Materials:} A gold coin or a tempting offer.
\textbf{Effect:} +1 die to Persuasion/Deceive involving deals \emph{or} sense when someone contemplates a dubious bargain.
\textbf{Push It:} Name a small inducement that counts as leverage, but the Red Prince takes note of the deal and may demand a favor in return.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Opened Palm (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Offering a gift with obvious strings attached.
\textbf{Effect:} When you give someone something they explicitly desire, they suffer −1 die to resist your next social action against them. The desire can be for information, an object, or even a feeling.
\textbf{Push It:} The target feels a compulsion to offer you something of "equal value" in return, but the compulsion affects their judgment and may lead them into dangerous situations.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Seal with Infernal Ink (Standard, 6 XP)} \emph{Action; Self; No.}
\textbf{Materials:} A contract signed in blood or ash.
\textbf{Effect:} Choose: +2 effect on one action \emph{or} start \emph{Controlled} on one action; then mark a fitting Condition (e.g., \emph{Compromised/Indebted/Tempted}). \emph{Or} downgrade a \emph{Miss}→\emph{Partial} or \emph{Partial}→\emph{Success \& Cost}.
\textbf{Push It:} The target owes a small favor, but the contract creates an equal but opposite obligation that must be fulfilled.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Calculated Debt (Standard, 7 XP)} \emph{Duration: Scene; Range: Near; Stacking: No.}
\textbf{Materials:} Pointing out a past kindness or injury.
\textbf{Effect:} You make a target acutely aware of a debt they owe (to you or someone else). They suffer −1 die to all actions until they take a concrete step to repay it, or until they perform a significant favor for you to "clear the slate."
\textbf{Push It:} The debt feels so pressing that the target must address it at the worst possible time, but the pressure affects their mental state and they may make reckless decisions.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Summon the Red Prince's Contract (High, 10 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A scroll of vellum that is never blank.
\textbf{Effect:} Bind one NPC to a significant agreement or suffer immediate supernatural consequence; \emph{or} +3 dice to all deal-making and one \emph{Miss→Success \& Cost} on social this scene.
\textbf{Push It:} Seed a clause that will return as a future complication, but the Red Prince's attention is drawn and he may demand additional concessions.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Inescapable Clause (High, 11 XP)} \emph{Duration: Permanent; Range: Touch; Stacking: No.}
\textbf{Materials:} The final signature on a magically binding pact.
\textbf{Effect:} You forge a supernatural contract with a willing or coerced signatory. The terms are enforced by the Red Prince's power. Breaking the contract results in a severe, predefined supernatural punishment (e.g., loss of a sense, a major curse, servitude). The GM has final veto on game-breaking terms.
\textbf{Push It:} The contract includes a hidden "fine print" that benefits you in a way the signatory did not anticipate, but the Red Prince takes a special interest in the contract and may interfere to ensure its fulfillment.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{The Traveler (Ways Between \& Roads)}
\paragraph{Set the Waystone (Low, 5 XP)} \emph{Duration: Scene; Range: Self; Stacking: Yes (travel only).}
\textbf{Materials:} A chalk sigil or ribboned pebble placed by the Thiasos at a threshold, milestone, or crossroads.
\textbf{Effect:} Start \emph{Controlled} on one Traverse/Scout movement this scene \emph{or} treat one terrain/visibility penalty as one step less severe.
\textbf{Push It:} One ally in \emph{Close} also gains the same benefit on their next movement, but the waystone marks your passage and may attract unwanted attention.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Quickened Step (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A light pack and comfortable shoes.
\textbf{Effect:} You move with preternatural ease. Ignore fatigue from a long walk. +1 die to rolls to keep pace with a faster traveler or to navigate a crowded street.
\textbf{Push It:} You can take a "shortcut," allowing you to arrive at a nearby destination a few minutes ahead of expected time, but the shortcut may lead through unexpected territory.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Walk the Hidden Mile (Standard, 9 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A thread measured to your stride, knotted by the Thiasos, and a sketched route.
\textbf{Effect:} Once per beat, if a plausible path exists, you may move \emph{one additional zone} as part of a successful action; you may also ignore engagement \emph{once} this scene while repositioning. Alternatively, when a travel/pursuit complication hits you, shift it from \emph{Spades/Diamonds} (harm/hazard) to \emph{Clubs} (delay/resources).
\textbf{Push It:} Bring one ally in \emph{Close} along when you take your bonus zone shift, but the hidden path affects both of you and may lead to unexpected destinations.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Guided Path (Standard, 8 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A map you've annotated with personal notes.
\textbf{Effect:} You intuitively sense the safest or most discreet route to a known destination within the current region. You and your group cannot become lost, and you gain +1 effect to avoid random encounters or patrols while traveling.
\textbf{Push It:} The path reveals a single, hidden point of interest along the way, but the revelation comes with a price that must be paid.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Unfold the Pilgrim's Road (High, 16 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} Dust from a crossroads traced into a narrow arch, a road-nail or ring of keys held by the Thiasos.
\textbf{Effect:} Choose one:
\emph{(A)} Open a short-lived \emph{Way-Between} linking two \textbf{known thresholds} within the current map; you and up to \emph{Near} allies may traverse it instantly once, ignoring engagement; \emph{or}
\emph{(B)} For this scene, treat your first zone change each beat as \emph{free}, and gain +2 dice on chase/escape/navigation actions.
\textbf{Push It:} Either extend the portal to include all allies in \emph{Near} \emph{or} force current pursuers to lose one position step, but the Way-Between becomes partially permanent and may attract unwanted traffic (generate 2 CP).
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 6 segments.

\paragraph{Rite of the Endless Road (High, 15 XP)} \emph{Duration: Extended; Range: Self; Stacking: No.}
\textbf{Materials:} A journey of a hundred miles begun with a single step.
\textbf{Effect:} You and your companions undertake a long journey. The travel time is halved, and the group arrives refreshed, clearing all minor Fatigue related to travel. Additionally, you may ask the GM one far-reaching question about the destination or what you might encounter on the way, receiving a vague but truthful omen.
\textbf{Push It:} The journey also provides a crucial piece of information or a minor beneficial item relevant to your goal at the destination, but the road remembers your passage and may exact a toll later.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{The Carrion-King (Decay \& Cycles)}
\paragraph{Rite of Gentle Rot (Low, 5 XP)} \emph{Duration: Instant; Range: Touch; Stacking: Yes (decay only).}
\textbf{Materials:} A piece of spoiled food or a dead insect.
\textbf{Effect:} Accelerate natural decay on one small, non-living object (e.g., rot a rope, tarnish a metal lock, spoil a meal). This grants +1 effect to actions like "Break" or "Sabotage."
\textbf{Push It:} The decay spreads slightly, affecting a second, similar object in Close range, but the accelerated decay attracts scavengers and vermin.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Wilting Bloom (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A withered flower.
\textbf{Effect:} You exude an aura of mild decay. You gain +1 die to resist disease and poison. Food you carry is preserved from further rot, but becomes tasteless.
\textbf{Push It:} A single small plant or source of food within Near range withers to dust, but the decay affects your own vitality and you suffer Fatigue 1.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Cycle's Turn (Standard, 8 XP)} \emph{Duration: Scene; Range: Touch; Stacking: No.}
\textbf{Materials:} A recently deceased creature (dead less than an hour).
\textbf{Effect:} Draw vitality from death to fuel creation. Choose one: Purify a small amount of food/water; cause a patch of useful fungi/herbs to sprout; or grant yourself or an ally +1d on the next roll to resist disease or poison.
\textbf{Push It:} The created substance is of exceptional quality or potency, but the cycle's turn affects the balance of life and death in the area.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Peaceful Rest (Standard, 7 XP)} \emph{Duration: Instant; Range: Near; Stacking: No.}
\textbf{Materials:} A handful of grave dirt sprinkled over a corpse.
\textbf{Effect:} Lay a minor spirit to rest. This prevents a corpse from being easily animated as undead, or quiets a minor haunting, granting +2 dice to the next social roll with any remaining spirits or mourners.
\textbf{Push It:} The soothing effect grants you or an ally temporary immunity to fear effects from undead for the scene, but the peaceful rest comes at the cost of disturbing other spirits.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Final Compost (High, 13 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} A handful of grave dirt.
\textbf{Effect:} Accelerate decay within a zone. Structures become unstable (−1d to rolls relying on them), and opponents suffer −1d on rolls to maintain equipment or complex efforts. Alternatively, you can rapidly consume a large obstacle (a barricade, a thick door) over the course of the scene.
\textbf{Push It:} The area becomes choked with fertile but sickening growth, inflicting a \emph{Sickened} condition on those who remain in the zone for more than a few beats, but the growth may prove useful later.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Great Cycle (High, 14 XP)} \emph{Duration: Extended; Range: Touch; Stacking: No.}
\textbf{Materials:} Burying a seed in rich, rotten earth.
\textbf{Effect:} You can ritually transform a significant amount of dead matter (a large corpse, a fallen tree) into something new over an extended period. This could be a fertile patch of land, a unique alchemical ingredient, or a temporary, beneficial environmental feature. The GM sets the scale and time required.
\textbf{Push It:} The process is accelerated, taking only a single scene, but is more visually dramatic and obvious, attracting attention from those who would interfere with the cycle.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{The Gallow's Bell (Consequences \& Retribution)}
\paragraph{Rite of the Whispered Name (Low, 4 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} A noose fragment; whispered confession.
\textbf{Effect:} Mark a target for "attention" - GM must introduce a complication related to their recent actions within 3 scenes.
\textbf{Push It:} Complication occurs immediately in current scene, but the whispered name echoes and may affect others who share the target's name.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Curse of the Unsettled Sleep (Low, 5 XP)} \emph{Duration: Scene; Range: Self; Stacking: No.}
\textbf{Materials:} Dream-catcher woven with guilt; midnight vigil.
\textbf{Effect:} Target dreams of their misdeeds, creating roleplay opportunities and narrative weight.
\textbf{Push It:} Dreams provide a true omen about future consequences, but the dreams also reveal one of your own secrets to the target.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Broken Mirror (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Shattered mirror; drop of the target's blood.
\textbf{Effect:} Target's reflection shows their "true nature" to others - social actions create narrative complications when deception attempted.
\textbf{Push It:} Reflection becomes actively hostile, creating social tension, but the mirror's magic affects all reflections in the area.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Mark of the Unfinished Business (Standard, 7 XP)} \emph{Duration: Extended; Range: Touch; Stacking: No.}
\textbf{Materials:} Unfinished letter; broken chain link.
\textbf{Effect:} Target feels narrative weight of incomplete obligations - story complications until resolution.
\textbf{Push It:} The unfinished business becomes more specific and demanding, but you become partially responsible for helping resolve it.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Curse of the Singing Chain (High, 13 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Iron chain that held a prisoner; funeral bell hammer.
\textbf{Effect:} Target hears echoes of every promise they've broken - suffer Fatigue 1 and narrative social complications when lying.
\textbf{Push It:} Chain manifests physically, restricting movement narratively, but the chain's magic affects all chains in the area.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{The Bell That Rings Thrice (High, 14 XP)} \emph{Extended; Zone; No.}
\textbf{Materials:} Gallows rope; three iron bells rung in sequence.
\textbf{Effect:} Target suffers three escalating narrative consequences related to their nature/sins over the next session.
\textbf{Push It:} Consequences affect allies/family members as well, but the bell's toll marks you as an agent of retribution.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation: 7 segments.}

\subsection{The Sealed Gate (Banishment, Wards \& Seals)}
\paragraph{Rite of the Marked Boundary (Low, 4 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}
\textbf{Materials:} Chalk mixed with salt; drawn sigil at threshold.
\textbf{Effect:} +2 dice to resist entry by marked supernatural threat type; create visible ward that deters 1-2 CP worth of minor entities.
\textbf{Push It:} Ward also alerts you when triggered by a mental ping, but the ward's magic attracts the attention of other supernatural entities.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Seal of the Forbidden Door (Low, 5 XP)} \emph{Duration: Extended; Range: Touch; Stacking: No.}
\textbf{Materials:} Seven locks of different metals; blessed chain.
\textbf{Effect:} Seal a door, chest, or passage against unauthorized entry - any attempt triggers 2 CP of supernatural effects.
\textbf{Push It:} Seal becomes permanent until properly broken with ritual, but the seal's magic affects all locks and seals in the area.
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}
\paragraph{Rite of the Cleansing Fire (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Holy oil; silver brazier inscribed with banishment runes.
\textbf{Effect:} Banish minor supernatural presence from area - force 1-2 CP worth of entities to retreat or suffer -2 dice to actions.
\textbf{Push It:} Banishment becomes permanent for the scene, but the cleansing fire affects all supernatural entities, friend or foe.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Ward of the Unbroken Circle (Standard, 7 XP)} \emph{Duration: Extended; Range: Zone; Stacking: No.}
\textbf{Materials:} Continuous loop of iron wire; seven candles of different colors.
\textbf{Effect:} Create a ward that prevents teleportation/planar travel within zone - all such attempts generate 2 CP of backlash.
\textbf{Push It:} Ward also blocks summoned entities from entering, but the ward's magic affects all teleportation and planar travel, including your own.
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}
\paragraph{Command of the Sealed Realm (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Lead box containing name of realm; key made from meteoric iron.
\textbf{Effect:} Seal an entire zone against supernatural entry - all entities of specified type suffer Desperate position to enter.
\textbf{Push It:} Seal extends to block planar travel in/out of zone, but the seal's magic affects all planar travel and may trap you within the zone.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Final Banishment (High, 14 XP)} \emph{Instant; Near; Stacking: No.}
\textbf{Materials:} Vessel containing the entity's true name; silver mirror facing away.
\textbf{Effect:} Permanently banish a single supernatural entity to its native realm - it cannot return without equal magic.
\textbf{Push It:} Banishment damages the entity, imposing permanent -2 dice to all its actions, but the banishment's magic affects the fabric of reality and may create unintended consequences.
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}
\emph{Obligation:} 7 segments.

\subsection{The Oath of Flame and Light (Radiance, Vows, and Holy Fire)}
\paragraph{Lay on Hands of Ember (Low, 6 XP)} \emph{Duration: Action; Range: Touch; Stacking: Yes (healing/warding only).}  
\textbf{Materials:} The Symbol warmed over a true flame while the Familiar intones a vow of care.  
\textbf{Effect:} Choose one: \textbf{(a)} \emph{Step down} one level of \emph{mundane or shadow-touched Harm} on the target, \textbf{or} \textbf{(b)} \emph{clear 1 Fatigue} from the target, \textbf{or} \textbf{(c)} grant the target \emph{start Controlled} on their next Resist vs fear/corruption this scene.  
\textbf{Push It:} Also remove one minor lingering condition tied to dread or pain (\emph{Shaken}, \emph{Rattled}, etc.), but you take on a portion of the target's suffering.
\textbf{Requires:} Familiar + Patron's Symbol. \;\; \textbf{Invoke:} 1 Boon. \;\; \textbf{Obligation:} \emph{5 segments.}

\paragraph{Brand of the Dawn (Standard, 9 XP)} \emph{Duration: Action; Range: Near (mark), Self (strike); Stacking: No.}  
\textbf{Materials:} Trace the sun-sign in the air with the Symbol; the Familiar draws a spark across your blade or palm.  
\textbf{Effect:} Mark one undead, demon, or outsider you can see. Until the end of the scene, your next successful strike against the marked target gains \textbf{+2 effect}. On that hit you may declare \emph{Turn}: the target must \emph{withdraw one zone immediately} or \emph{stand their ground and generate 2 CP (Spades)}.  
\emph{Alternatively,} inscribe the sun-sign on a threshold: for this scene, named foes crossing it suffer \textbf{-1 die} on their next action.  
\textbf{Push It:} When you declare \emph{Turn}, extend the choice to \emph{one additional} engaged foe of the same type, but the brand's light attracts other undead or demons to the area (generate 2 CP).
\textbf{Requires:} Familiar + Codex + Patron's Symbol. \;\; \textbf{Invoke:} 1 Boon. \;\; \textbf{Obligation:} \emph{5 segments.}

\paragraph{Sanctuary of the Just Flame (High, 13 XP)} \emph{Duration: Scene; Range: Zone; Stacking: No.}  
\textbf{Materials:} A ring of salt and light traced with the Symbol while the Familiar recites your oath by name.  
\textbf{Effect:} Consecrate the zone. Undead, demons, and outsiders that \emph{enter or act} within the sanctuary suffer \textbf{-1 die} on their first action each beat. Allies within the zone \emph{immediately clear 1 Fatigue} and gain \textbf{+1 die} to Resist fear/corruption for the scene. \emph{Once this scene,} name one such foe in the zone: they must \emph{withdraw from the sanctuary immediately} or suffer a \textbf{significant immediate consequence} (GM spend, typically \emph{Spades} or \emph{Hearts}).  
\textbf{Push It:} Extend the sanctuary's edge to an adjacent zone for one beat \emph{or} attempt a decisive banishment against a single marked foe (test \emph{Spirit + Faith}, DV by foe's might; on success they are expelled from the scene), but the sanctuary's power draws the attention of powerful undead or demonic entities.
\textbf{Requires:} Familiar + Codex + Tier III + Patron's Symbol. \;\; \textbf{Invoke:} \textbf{2 Boons}. \;\; \textbf{Obligation:} \emph{7 segments.}

\section{Patron Relations and Jealousy}

Patrons do not share their chosen ones willingly. When a character bound to one Patron attempts to invoke rites of another, the first Patron's jealousy manifests as a complication (typically 1 CP) that reflects their domain and personality. This penalty does not apply to characters following the Invoker path, as they maintain no exclusive Patron relationships.

\subsection{Patron Relationship Dynamics}

\begin{table}[htbp]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Primary Patron} & \textbf{Relationship} & \textbf{Secondary Patron} & \textbf{Jealousy Response} \\
\hline
Any & Favor & Any Favored Patron & Never Jealous \\
\hline
Any & Loath & Any Loathed Patron & Always Jealous (+1 Obligation) \\
\hline
Any & Ambivalent & Any Ambivalent Patron & Sometimes Jealous (50\% chance, +1 Obligation) \\
\hline
\end{tabular}
\caption{Patron Jealousy Response Matrix}
\end{table}

\subsection{Key Patron Relationships}

\subsubsection{Major Conflicts}
\begin{itemize}
\item \textbf{Khemesh ↔ Raéyn}: Loath each other (depths vs. tides)
\item \textbf{Mor'iraath ↔ Mykkiel}: Loath each other (chaos vs. law)
\item \textbf{The Witness ↔ Maelstreus}: Loath (truth vs. deception)
\item \textbf{Ikasha ↔ Mykkiel}: Loath (shadow vs. law)
\item \textbf{Inaea ↔ Isoka}: Loath (rigid fate vs. fluid change)
\item \textbf{The Sealed Gate ↔ The Outsider}: Loath (containment vs. intrusion)
\item \textbf{The Carrion-King ↔ The Gallow's Bell}: Loath (natural decay vs. imposed judgment)
\item \textbf{Oath of Flame and Light ↔ Mor'iraath}: Loath (purity vs. corruption)
\item \textbf{Oath of Flame and Light ↔ The Outsider}: Loath (sacred order vs. chaos)
\end{itemize}

\subsubsection{Strong Alliances}
\begin{itemize}
\item \textbf{Ikasha → Raéyn}: Favor (shadow flows with tides)
\item \textbf{Isoka → Mor'iraath}: Favor (change feeds consumption)
\item \textbf{Inaea → Maelstreus}: Favor (webs connect to deals)
\item \textbf{The Traveler → Raéyn}: Favor (movement respects flux)
\item \textbf{Khemesh → The Carrion-King}: Favor (both embrace decay and depth)
\item \textbf{Varnek Karn → The Carrion-King}: Favor (both death-as-process)
\item \textbf{The Gallow's Bell → Mykkiel}: Favor (both concerned with justice/judgment)
\item \textbf{Oath of Flame and Light → Mykkiel}: Favor (both concerned with law and order)
\item \textbf{Oath of Flame and Light → The Gallow's Bell}: Favor (both concerned with justice and consequences)
\item \textbf{Oath of Flame and Light → The Sealed Gate}: Favor (both concerned with protection and sanctity)
\end{itemize}

\subsubsection{Complex Relationships}
\begin{itemize}
\item \textbf{Sacred Geometry ↔ Clockwork Demon}: Ambivalent (order vs. chaos, but both systematic)
\item \textbf{Varnek Karn ↔ The Carrion-King}: Ambivalent (both death, different aspects)
\item \textbf{Nidhoggr ↔ Valewood}: Ambivalent (both ancient, different memories)
\item \textbf{The Traveler ↔ The Ways Between}: Ambivalent (both concerned with paths, but one is a patron, one a realm)
\item \textbf{Mab ↔ The Outsider}: Ambivalent (both exist in spaces between, but Mab is court-structured while Outsider is chaotic)
\item \textbf{Sacred Geometry ↔ The Sealed Gate}: Ambivalent (both concerned with boundaries, but one creates order, one enforces barriers)
\item \textbf{Nidhoggr ↔ The Gallow's Bell}: Ambivalent (both concerned with the weight of the past, but one preserves ancient wisdom while other enforces consequences)
\item \textbf{Oath of Flame and Light ↔ Ikasha}: Ambivalent (light vs. shadow, but both concerned with truth - revealed vs. hidden)
\item \textbf{Oath of Flame and Light ↔ The Witness}: Ambivalent (both concerned with truth, but through different means - revelation vs. observation)
\end{itemize}

\subsection{Patron Jealousy Manifestations}

\subsubsection{Immediate Effects}
When jealousy is triggered:
\begin{itemize}
\item \textbf{+1 Obligation Segment} on current patron's clock
\item \textbf{Thiasos Distress}: Familiar shows agitation (-1 die assistance)
\item \textbf{Ritual Interference}: -1 die to rites involving offending patron
\item \textbf{GM Complication}: 1-2 CP of immediate narrative consequences
\end{itemize}

\subsubsection{Escalating Manifestations}
When Obligation Clock fills due to jealousy:

\paragraph{The Witness}
\begin{itemize}
\item Surveillance by unknown observers
\item Mysterious documents revealing duplicity
\item Whispers about secrets being shared
\end{itemize}

\paragraph{Sacred Geometry}
\begin{itemize}
\item Mathematical anomalies in rituals
\item Geometric patterns appearing accusingly
\item Structural instabilities in workspace
\end{itemize}

\paragraph{Ikasha}
\begin{itemize}
\item Shadows showing disapproval
\item Dreams of being watched
\item Whispers of betrayal in darkness
\end{itemize}

\paragraph{Khemesh}
\begin{itemize}
\item Oppressive weight sensations
\item Salt water appearing in wrong places
\item Deep rumbling sounds
\end{itemize}

\paragraph{Raéyn}
\begin{itemize}
\item Contrary winds/tides
\item Unpredictable weather
\item Messages containing warnings
\end{itemize}

\paragraph{Clockwork Demon}
\begin{itemize}
\item Device malfunctions during rituals
\item Gears grinding ominously
\item Steam hissing disapprovingly
\end{itemize}

\paragraph{Isoka}
\begin{itemize}
\item Unwanted physical transformations
\item Items changing unexpectedly
\item Paths shifting to complicate travel
\end{itemize}

\paragraph{Inaea}
\begin{itemize}
\item Coincidences with personal significance
\item Web patterns appearing accusingly
\item Strings tangling at crucial moments
\end{itemize}

\paragraph{Mor'iraath}
\begin{itemize}
\item Uncontrolled flames in rituals
\item Items heating uncomfortably
\item Smoke forming accusatory shapes
\end{itemize}

\paragraph{Mykkiel}
\begin{itemize}
\item Legal challenges to authority
\item Oaths requiring complex reaffirmation
\item Officials questioning credentials
\end{itemize}

\paragraph{Rite of the Borrowed Mantle (Advanced, 18 XP)} \emph{Duration: Instant; Range: Self; Stacking: No.}
\textbf{Materials:} A personal sacrifice worthy of the Patron (GM defined).
\textbf{Effect:} Before a roll, declare you are borrowing your Patron's mantle. You manifest a minor physical aspect. You may upgrade the effect level of your action by two steps (e.g., from Limited to Great) or add +3 dice to the roll.
\textbf{Push It:} The manifestation is so potent it affects the scene, creating a temporary environmental aspect related to your Patron that you can leverage once.
\emph{Requires: Familiar + Codex + Tier III + Special GM Approval \ (\textit{Invoke:} \textbf{3 Boons} and advance the Obligation Clock by 2 segments).}

\paragraph{Favor \& Attention}
\emph{Flavor.} A personified current of righteous light and purifying flame. It answers to oaths kept, mercy shown, and courage under shadow.  
\emph{Favor Suits.} \textbf{Hearts} (zeal, conscience), \textbf{Diamonds} (revelation, radiant signs), \textbf{Spades} (searing backlash).  
\emph{Patron's Symbol (examples).} A sun-sign stamped in cold iron; a glass ampoule holding a never-quenched ember; a burnished vow-ring engraved with a sunrise. You must possess the Symbol keyed to this Patron to Invoke its Rites.

\paragraph{Rite Tags}
Unless noted otherwise, each Rite follows the core procedure: \textbf{Invoke} (spend 1 Boon; mark +1 Obligation); once per scene the \textbf{Push It} rider may add +1 segment; a \textbf{Backlash} Miss or Success \& Cost on an aided roll adds +1 segment; max \textbf{Tier +3} segments/scene per Rite.

\paragraph{Access Ladder}
\emph{Low-Power} requires \textbf{Familiar + Patron's Symbol}. \;\;
\emph{Standard} requires \textbf{Familiar + Codex + Symbol}. \;\;
\emph{High-Power} requires \textbf{Familiar + Codex + Tier III + Symbol} and \textbf{Invoke 2 Boons}.

\bigskip

\section{Patron Manifestations}

When the Obligation Clock fills, a Patron will start to take interest in their charge. 

\begin{table}[htbp]
\centering
\begin{tabular}{>{\bfseries}p{0.8cm} p{4cm} p{8.5cm}}
\toprule
\textbf{d6} & \textbf{Manifestation Concept} & \textbf{Specific Patron Examples \& Lore Hooks} \\
\midrule
\textbf{1} & \textbf{The Subtle Shift} — Reality itself subtly realigns to reflect the Patron's nature. The change is environmental and persistent. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{The Witness:} Shadows hold unnaturally steady; mirrors show frozen flashes of the recent past. \\
        \textbf{The Sacred Geometry:} Angles drift toward pleasing ratios; spirals appear in wood grain and flowing water. \\
        \textbf{Oath of Flame and Light:} Candles within sight burn steady and bright; shadows retreat a finger's breadth from your skin; fresh wounds close into pale, clean scars overnight.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{2} & \textbf{The Unwelcome Gift} — The Patron bestows a power or insight that is useful but corrosive, demanding to be used. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{Ikasha:} You hear thoughts that hide in shadowed places; the whispers are accusatory and paranoid. \\
        \textbf{Mor'iraath:} Your touch smolders: you can kindle dry tinder at will, but delicate things you handle char and scar. \\
        \textbf{Oath of Flame and Light:} Your hands shed honest light: lies within arm's reach rasp and falter, yet small kindnesses you try to conceal blaze obvious to onlookers.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{3} & \textbf{The Hungry Messenger} — A creature or phenomenon aligned with the Patron dogs your steps, observing or making small demands. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{Varnek Karn:} An intelligent rat adopts you—bringing coins and trinkets, leaving neat arrangements of bones in your pack. \\
        \textbf{Mab:} A thorny vine sprouts where you bled; it blooms with weeping flowers that demand daily tending. \\
        \textbf{Khemesh:} Your footprints briefly fill with brackish water; gulls circle, and ropes creak where none are hung.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{4} & \textbf{The Reflected Flaw} — The Patron turns your own methods back on you, highlighting a hypocrisy or dependency. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{Mykkiel:} Casual promises bind literally; "a minute" becomes a geas until fulfilled. \\
        \textbf{Isoka:} A mask you wear becomes your public face; strangers insist the caricature is who you are. \\
        \textbf{Inaea:} A small deception you set in motion weeks ago snaps taut and entangles you in its fallout.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{5} & \textbf{The Echoing Demand} — The Patron requires a tangible task or sacrifice related to their domain. This is a direct quest. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{The Traveler:} Do not sleep under the same roof twice for a week, or dream of a road erasing your hometown from its maps. \\
        \textbf{Nidhoggr:} Recover and preserve a forgotten piece of history or suffer suffocating visions of being buried in time. \\
        \textbf{Maelstreus:} Broker a specific, morally gray bargain within a tenday or watch a profitable pact of yours unravel.
    \end{tabular} \\
\cmidrule(lr){2-3}
\textbf{6} & \textbf{The Personal Visitation} — The Patron, or a powerful emissary, appears in a dream or liminal space to deliver a warning or pivotal choice. &
    \begin{tabular}[t]{p{8.5cm}}
        \textbf{The Outsider:} At the third hour before dawn, an extra, impossible corner opens in your room; a silent figure offers a key to a door that should not exist. \\
        \textbf{Oath of Flame and Light:} During a vigil, a figure of dawnfire offers to seal a vow: accept and gain strength to banish one named trespass, but relinquish a cherished grudge or secret. \\
        \textbf{Mab:} You attend a dream-feast where words are binding; break the table's rules and the consequences follow you into morning.
    \end{tabular} \\
\bottomrule
\end{tabular}
\end{table}

\section*{GM Guidance for Using the Table}

\begin{enumerate}
    \item \textbf{Weave, Don't Announce:} The manifestation should be introduced subtly. Don't say, ``The Witness is angry.'' Instead, describe how the PC notices that their shadow is moving a half-second too slow.

    \item \textbf{Start a Clock:} Each manifestation is the first segment of a new, typically 4-segment ``Patron's Attention'' clock. If this new clock fills, the consequences escalate from inconvenient to dangerous.

    \item \textbf{Offer a Remedy:} The manifestation should suggest a way to ``pay the debt'' or reduce the clock. For example, to appease \textbf{Varnek Karn}, the PC might need to give the rat a proper burial for a forgotten soul. Succeeding clears one segment.

    \item \textbf{Tie to the PC's Actions:} The chosen manifestation should feel like a direct consequence of how the PC has been using the Rites. A PC who used \textbf{Mor'iraath's} rites for destruction might get ``The Hungry Messenger'' (a small fire elemental that eats their belongings), while one who used them for passion might get ``The Reflected Flaw'' (their emotions become dangerously contagious).
\end{enumerate}

\section*{Imbunement Rites}

\begin{quote}\itshape
``He says the Gate wants the city sealed. I say it wants it opened---so that it may \emph{shut properly} when the time comes. The Gate smiles at us both.''
\end{quote}
\begin{flushright}
--- Scribe-Advocates of the Sealed Threshold, \textit{Third Petition}
\end{flushright}

\subsection*{Patron's Gift}
\textbf{Patron's Gift} (Free, Requires Thiasos )\\
Duration: Scene; Range: Touch; Stacking: No.\\
Materials: Personal item of significance.\\
Effect: Imbue one item with temporary magical power related to your Patron's domain. The item functions as a magical weapon(+1 Melee) and specialized tool (See Below) for this scene with +1 effect on relevant actions.\\
Push It: The item's power persists for one additional scene but marks +1 segment on your Obligation Clock.\\
Requires: Familiar (Invoke: 1 Boon).

\subsection*{Patron Imbunements}

\begin{table}[htbp]
\centering
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{>{\bfseries}p{3.2cm} >{\raggedright\arraybackslash}p{6.1cm} >{\raggedright\arraybackslash}p{6.1cm}}
\toprule
\textbf{Patron} & \textbf{Basic Effect} & \textbf{Push It Effect} \\
\midrule
The Witness & Item bears a truth-mark — \textbf{+1 die} to \emph{Investigation/Notice} & Item reveals one hidden truth about the current target or environment \\
The Sacred Geometry & Item settles into perfect ratio — \textbf{+1 die} to precision, timing, or measured tasks & Item ignores the first penalty from terrain or unstable footing on one action \\
Ikasha & Item takes on umbral aspect — \textbf{+1 die} to \emph{Stealth/Deceive} & Item (or its wielder's hand) may pass through one solid barrier unimpeded (once) \\
Khemesh & Item gains crushing undertow — \textbf{+1 effect} on raw force or break actions & On one hit, item ignores armor/hardness \\
Raéyn & Item flows with the turn — \textbf{+1 die} to adaptation, navigation, or weathered travel & Treat one environmental hazard or shift as one step less severe (once) \\
The Clockwork Demon & Item ticks with proscribed precision — \textbf{+1 die} to \emph{Craft/Repair} & Item functions despite one missing or broken component (once) \\
Isoka & Item loosens bonds — \textbf{+1 die} to \emph{Escape/Transform} actions & Shed one minor physical or mental condition (once) \\
Inaea & Item threads the social weave — \textbf{+1 die} to \emph{Influence/Manipulate} & Expose one relevant connection, obligation, or leverage between NPCs \\
Mor'iraath & Item wreathed in hungry flame — \textbf{+1 effect} on destructive actions & Flammables ignite on contact with a successful hit (this scene) \\
Mykkiel & Item bears the seal — \textbf{+1 die} to \emph{Authority/Command} & Compel minor, immediate compliance from a lawful subject (once) \\
Varnek Karn & Item tastes of the barrow — \textbf{+1 effect} vs. living targets & Ignore 1 \emph{Fatigue} on a single exertion (once) \\
Mab & Item is fae-touched — \textbf{+1 die} to \emph{Charm/Deceive} & Conjure a minor illusion or misdirection (once) \\
Nidhoggr & Item hums with antiquity — \textbf{+1 die} vs. ancient or supernatural wards & Reveal one pertinent historical fact or true name about the target \\
The Outsider & Item phases at the edge — \textbf{+1 die} vs. material barriers or confinement & Briefly exist "between" to bypass one obstacle or occupy two points for an instant (once) \\
Maelstreus & Item tempts the ledger — \textbf{+1 effect} on gambles, bargains, or risky leverage & Convert one \emph{Miss} into \emph{Success \& Cost} (once) \\
The Traveler & Item keeps the way — \textbf{+1 die} to travel, navigation, or pursuit/escape & Treat one difficult-terrain or engagement penalty as null for a single movement \\
The Carrion-King & Item hastens decay — \textbf{+1 effect} on sabotage or corrosion & Cause one organic object to rot, split, or foul beyond use (once) \\
Oath of Flame and Light & Item bears dawnfire — \textbf{+1 die} to \emph{defend/heal} or when striking undead, demons, or outsiders & On a successful hit against such a foe, \emph{Turn} it (it withdraws, DV=Tier) \emph{or} it immediately generates 2 CP (\emph{Spades}) \\
\bottomrule
\end{tabular}
\end{table}

\section{Ritual of Severance: Exorcism of Patronage}
\label{subsec:exorcism}
\textbf{Name:} \emph{Cut the Knot of Patronage} \quad \textbf{Type:} Ritual (Extended) \\
\textbf{Who:} The warlock, \emph{plus} 1--3 assistants (any Arts; distinct procedures). \\
\textbf{Materials:} A mirror veiled in salt-thread, anointed Codex page or Thiasos token, a personal sacrifice tied to the Patron's domain. \\
\textbf{Procedure:} Over an extended scene, the circle denies \emph{names, signs, and vectors} that bind the Patron. Each helper offers a distinct symbolic inversion tied to their skill.
\begin{itemize}
  \item \textbf{Cost:} The primary invoker spends \textbf{2 Boons}. Mark \emph{Exposure +1}. Each helper risks 1 Fatigue on a \emph{Miss}.
  \item \textbf{Test:} GM sets DV by entanglement (typ. DV 2–3). On success, \textbf{clear the Rite's Obligation Clock} to 0 and \textbf{choose one}: (a) \emph{Dismiss} your Thiasos \underline{or} (b) \emph{Seal} your Codex (cannot be used for that Patron until re-consecrated in Downtime).
  \item \textbf{On Success \& Cost:} As above, but the GM immediately starts a fading \emph{Echo} thread keyed to the Patron (2–3 CP over the next scene).
  \item \textbf{On Miss:} The knot tightens. Mark \emph{+2 segments} on the active Patron clock, and the GM introduces a \emph{Geas}-like demand or intrusion.
\end{itemize}
\textbf{Aftermath:} You may switch paths (Codex $\leftrightarrow$ Thiasos) once the severance is complete and the inert resource is relinquished or reconsecrated in Downtime.


\section*{Invoker Path: Ritual Access}

Instead of the Familiar + Codex approach, an Invoker accesses Rites through \textbf{Symbol + known Rite}. All Invoker Rites are \textbf{rituals} and take \textbf{Significant time} (typically 10–30 minutes; GM sets fiction). Invoker Rites \emph{do not} use Push It. The caster must openly display the relevant \textbf{Patron's Symbol} throughout.

\paragraph{Attention \& Obligation (Invoker)}
Completing an Invoker ritual \textbf{marks +1 segment} on that Rite's \emph{Obligation Clock} (even if outside a scene). This replaces any Boon cost. If you \emph{Crack the Seal} (see below) to act instantly, mark more.

\subsubsection*{Ritual Novice (4 XP)}
\textbf{Effect:} You may perform any \textbf{Low-Power} Rite you \emph{know} as a sacred ritual. The Rite's duration becomes \textbf{Extended} unless it is already longer.\\
\textbf{Limitations:} No Push It; suitable environment; visible Symbol throughout.\\
\textbf{Crack the Seal (Instant Cast):} Immediately resolve the Rite without time. \textbf{Set the Symbol to \emph{Compromised}} and \textbf{mark +2 segments} on that Rite's Obligation. The GM may immediately spend \textbf{1 CP} related to the Patron's domain.\\
\textbf{Jealousy:} Invoking a Patron you are not pledged to \textbf{also} marks \textbf{+1 segment} on that Patron's \emph{Obligation}.

\subsubsection*{Ritual Adept (6 XP)} 
\textbf{Effect:} As above, and you may perform \textbf{Standard} Rites you \emph{know}.\\
\textbf{Requires:} Ritual Novice.

\subsubsection*{Ritual Master (8 XP)} 
\textbf{Effect:} As above, and you may perform \textbf{High-Power} Rites you \emph{know}.\\
\textbf{Instant Cast (High-Power):} When you \emph{Crack the Seal} on a High-Power Rite, \textbf{mark +3 segments} on that Rite's Obligation and set the Symbol to \emph{Compromised}.\\
\textbf{Requires:} Ritual Adept.

\paragraph{Concurrent Bindings (cap)}
You may sustain at most \textbf{Spirit} ongoing Invoker rituals at once. Beginning a new one forces the oldest to lapse or immediately marks \textbf{+1 segment} on its Obligation (your choice).

\paragraph*{Symbol Vulnerability }

When a Patron's Symbol is compromised through damage, theft, or magical interference, it cannot be used for Rite invocation until repaired. Compromised Symbols:
\begin{itemize}
\item Cannot be used for 24 hours
\item Require Downtime repair (1 day) or XP (2 XP) to restore
\item Risk of permanent damage on 3+ compromises
\end{itemize}

The GM may introduce complications when a Symbol is compromised, such as attracting unwanted attention from entities aligned with that Patron or causing minor magical malfunctions in the bearer's other equipment.

\paragraph{Multi-Symbol Risk}
You may carry multiple Symbols (one per Patron). If you carry \textbf{3+} different Patrons' Symbols, the first Invoker ritual you complete each scene marks \textbf{+1 additional segment} on its Obligation (the choir gets loud).

\subsection*{Patron's Symbol (Minor Asset)}\index{Patron's Symbol}\index{Assets}

\textbf{Type:} Minor Asset (4 XP) \hfill \textbf{Scope:} One specific Patron

\textbf{Requirement.} To Invoke any Rite of a Patron via the Invoker path, you must \textbf{possess that Patron's Symbol}. You may carry only \textbf{one} Symbol per Patron (but may hold Symbols of different Patrons).

\textbf{What it is.} A consecrated token keyed to a Patron's office (sign, seal, relic). It anchors names and permissions.

\textbf{Condition \& Upkeep.}
\begin{itemize}
  \item \textbf{Maintained:} Functions normally.
  \item \textbf{Neglected:} Ritual completes but \textbf{mark +1 extra Obligation segment} on that Rite.
  \item \textbf{Compromised:} Inert—cannot be used until \textbf{restored}.
\end{itemize}

\textbf{Crack the Seal (Instant Cast).} As part of an Invoker Rite, you may \textbf{immediately} resolve the effect by \textbf{setting the Symbol to \emph{Compromised}} and marking \textbf{+2 segments} (\textbf{+3} if High-Power) on that Rite's Obligation. The GM may spend \textbf{1 CP} on-theme. \emph{This does not remove the Asset.}

\textbf{Restore a Symbol (Downtime).} Repair a \emph{Compromised} Symbol with a Downtime action and a fitting test (DV 3 or by fiction). On success, set to \emph{Maintained}; on a shaky result, it returns as \emph{Neglected}. Alternatively, spend \textbf{1 XP} to fully restore it between sessions.

\textbf{On-Screen Activation (once/scene).} Presenting/working the Symbol costs \textbf{1 Boon}: \textbf{+1 die} \emph{or} \textbf{improve Position by one step} for negotiation, warding, or parley with that Patron's agents/influences.

\textbf{Off-Screen (1/session).} Petition a sign, omen, or minor audience relevant to the Patron (GM sets limits).

\paragraph{Loss, Theft, and Forgeries}
A stolen/forged Symbol cannot Invoke your Rites, but can misdirect omens. A false Sign treated as real begins a short "False Sign" complication thread.

\begin{table}[htbp]
\centering
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{>{\bfseries}p{4.0cm} >{\raggedright\arraybackslash}p{10.8cm}}
\toprule
\textbf{Patron} & \textbf{Symbol Examples (choose or craft one)} \\
\midrule
The Witness & Obsidian eye-sigil pendant scored with a single lash; a silvered mirror shard wrapped in black wire; a wax seal-stamp bearing an open eye, carried on a cord. \\
The Sacred Geometry & Brass heptagram rule or compass with seven-point tooth; bone tablet inlaid with a golden-ratio spiral; plumb-bob knotted with a proof written in red thread. \\
Ikasha & Vial of soot-oil that drinks light; knot of black silk whose shadow forms a sigil; fingerbone ring lacquered matte, always cool to the touch. \\
Khemesh & Salt-etched iron chain link; lead sounder-weight engraved with abyssal curls; barnacle-bitten coin drilled and worn as a token. \\
Raéyn & Sea-glass disk notched for tides; storm-twined rope knot crusted with salt; glass ampoule of rainwater gathered at three crossings. \\
The Clockwork Demon & Gear tooth sealed in a tube of red oil; mainspring coil etched with forbidden numerals; rivet of unknown alloy stamped with a demon's stamp. \\
Isoka & Shedding of snakeskin braided into a wrist loop; quicksilver ampoule that gleams like a new skin; brass torque engraved with interlaced serpents. \\
Inaea & Spool of spider-silk thread around a bone bobbin; silver ring of interlocking knots; amber bead with a tiny inclusion "caught" at center. \\
Mor'iraath & Coal set in an iron claw; glass ampoule holding a live ember that never cools; charred vow-strip bound with copper wire. \\
Mykkiel & Cold-iron seal matrix engraved with a balance; square rule stamped with a code number; parchment writ-tag tied with red cord and wax. \\
Varnek Karn (The Carrion-King) & Ossuary bead rosary threaded with grave-dust; carved phalanx-bone tally; coin of bone and obsidian fused and drilled. \\
Mab & Hawthorn thorn wrapped in silver wire; mirror shard backed with green felt; acorn cup lined with red silk, worn as a charm. \\
Nidhoggr & Fossil tooth shard on a black thong; river-stone that has never seen sunlight, kept in a dark pouch; obsidian spindle with a hairline flaw. \\
The Outsider & Cracked mirror shard laced with silver wire; key with no teeth and no ward; crystal prism that throws an extra, impossible ray. \\
Maelstreus & Signet ring bearing infernal script; miniature contract scroll sealed in red glass; coin scored in halves and linked by a fine chain. \\
The Traveler & Road-nail wrapped in waythread; waystone pebble tied with ribbon and dusted in road-salt; brass compass missing its needle that nevertheless points. \\
Oath of Flame and Light & Cold-iron sun-stamp; glass ampoule with a consecrated spark; vow-ring engraved with a sunrise and the bearer's true name. \\
\bottomrule
\end{tabular}
\end{table}
\subsection*{Rite Access Ladders}

\paragraph{Invoker Path}
\begin{itemize}
  \item \textbf{Novice:} Low-Power Rites you know as rituals (\emph{+1 segment on completion})
  \item \textbf{Adept:} Standard Rites you know as rituals (\emph{+1 segment})
  \item \textbf{Master:} High-Power Rites you know as rituals (\emph{+1 segment})
  \item \textbf{Instant Cast (any tier):} \emph{Crack the Seal} (\emph{Symbol $\rightarrow$ Compromised}; \emph{+2 segments}, or \emph{+3} if High-Power; GM may spend 1 CP)
\end{itemize}

\subsubsection{Invoker Path Limitations}

The Invoker path trades immediacy and flexibility for reliability and accessibility. These limitations ensure that Invokers remain balanced against traditional Rite users while maintaining their unique identity as ritual specialists.

\paragraph{Casting Time Constraints} All Invoker Rites are \textbf{rituals} requiring \textbf{Significant Time} (typically 10-30 minutes). The GM sets the specific fiction requirements based on the Rite's nature and power level. Emergency casting is possible through the \textbf{Crack the Seal} mechanic, but at significant cost.

\paragraph{No Push It Benefits} Invoker Rites \textbf{cannot utilize Push It effects}. The ritualistic nature of Invoker magic requires precise, predetermined procedures that do not allow for the improvisational enhancements available to bonded Rite users.

\paragraph{Environmental Requirements} Rituals require a \textbf{suitable environment} appropriate to the Rite's nature. Complex rituals may require specific ceremonial spaces, ritual purity, or freedom from interruption. The GM determines environmental suitability based on fiction.

\paragraph{Symbol Display Requirement} The \textbf{Patron's Symbol must be openly displayed} throughout the ritual performance. Concealing or hiding the Symbol disrupts the ritual connection and prevents the Rite from functioning.

\paragraph{Concurrent Ritual Limit} An Invoker may sustain at most \textbf{Spirit} ongoing Invoker rituals simultaneously. Beginning a new ritual forces the oldest active ritual to lapse or immediately marks \textbf{+1 segment} on its Obligation Clock (caster's choice).

\paragraph{Multi-Symbol Overload} Carrying \textbf{4 or more different Patrons' Symbols} creates "choir" effects - the first Invoker ritual completed each scene marks \textbf{+1 additional segment} on its Obligation Clock as conflicting divine harmonics interfere with precise ritual work.

\paragraph{Attention & Obligation Generation} Completing an Invoker ritual marks \textbf{+1 segment} on that Rite's Obligation Clock \textbf{even outside scene context}. This represents the constant attention required to maintain ritual connections with powerful entities through mere symbols rather than bonded relationships.

\paragraph{Crack the Seal Consequences} Using the instant cast mechanism immediately sets the Symbol to \textbf{Compromised} and marks \textbf{+2 segments} (+3 for High-Power) on the Rite's Obligation Clock. The GM may immediately spend \textbf{1 CP} related to the Patron's domain theme, representing the instability of rushed ritual work.

\chapter{Closing Words}

\section*{The Price of Knowledge}

\lettrine[lines=3]{A}{}s I set down my pen and watch the ink dry on these final pages, I am struck by the weight of what I have recorded. Each ritual, each warning, each whispered secret represents not just power gained, but power that has been tested, measured, and found wanting. For knowledge without wisdom is merely dangerous, and power without purpose is merely destruction.

\shadow{Rukhir has been silent for hours now, his black eyes reflecting the candlelight like twin stars. Even he, ancient as he is, seems wearied by the cataloging of so much darkness.}

In my years of walking the shadow-paths, I have learned that the greatest magic is not in the spells we cast or the rituals we perform, but in the choices we make when the candles burn low and the whispers grow loud. Every ritual in this grimoire has been tested—not just for its efficacy, but for its cost. And make no mistake, dear reader, each one demands a price that would make merchants weep and kings tremble.

The Bonding Ceremony can unite hearts, but it can also chain souls. The Sanctuary Consecration can offer protection, but it can also become a prison. The Memory Reconstruction can reveal truth, but it can also unleash madness. Each path we walk in shadow leads both toward power and toward peril, and the wise witch knows that the difference lies not in avoiding the darkness, but in understanding it.

\section*{A Word to the Keeper}

If you have read this far, then perhaps you are ready for what lies ahead. But readiness is not the same as wisdom, and wisdom is not the same as safety. The rituals and \emph{Rites} I have shared with you are not toys for idle amusement or tools for petty revenge. They are conversations with powers who remember their due. Each Rite is a door you open with a \textbf{Symbol} or a carefully kept \textbf{Codex}, and every door opens \emph{both} ways.

\shadow{"Knowledge is a burden," Rukhir finally speaks, his voice like rustling leaves. "And every burden leaves marks upon the bearer."}

Use these workings with respect for those who taught me, for those who suffered for this knowledge, and for those who may yet suffer because of it. Shadow-magic is not evil in itself—nor are the Patrons simple saints or monsters. They are \emph{offices} of the world: Witness and Law, Tide and Road, Ember and Depth, the silk of Fate and the hiss of Change. In the hands of the unworthy, any Rite becomes a weapon; in the hands of the desperate, even a blessing can burn.

Remember always that Ikasha, She Who Sleeps Within Shadow, is not a goddess of chaos or evil. She is the patron of those who understand that truth often hides in darkness, that power comes at a cost, and that sometimes the only way to fight the monsters is to borrow their grammar. Others answer, too: the Geometry to proofs, the Traveler to thresholds, the Oath-bound Light to vows kept. They will \emph{listen}—and they will \emph{remember}.

\section*{The Path Forward}

For those who would continue down the shadow-roads, I offer these final pieces of advice:

First, never trust completely. Even the most faithful familiar has its own agenda, and even the most courteous Patron is jealous of attention. Trust, but verify. Love, but protect. Hope, but prepare. When you \emph{Invoke}, state your price and hear theirs.

Second, remember that power is a responsibility. The ability to influence others, to bend position, to peer into hidden places—these are not gifts to be used lightly. With each \emph{Rite} you perform, with each seal you break, you write your name deeper into the ledger of \emph{Obligation}. Keep your \textbf{Symbol} tended, your \textbf{Codex} clean, and your promises cleaner still.

Third, know when to walk away. There are secrets that should remain buried, powers that should remain unfound, and roads that should remain unopened. When Law and Tide argue, choose your shore. When the Witness offers truth that will unmake a friend, choose which world you will live in afterward. Wisdom is as much in knowing what not to do as in knowing what to do.

\shadow{Rukhir spreads his wings, and in the candlelight, they seem to hold fragments of every shadow I have ever seen. "The end of one path," he says, "is always the beginning of another."}

If the cords you have tied begin to choke you, there is a knife for knots: the severance rite. Use it without shame. Consent is a law deeper than any Patron's pride.

Finally, remember that you are not alone. The Tulkani have a saying: "No shadow falls without light to cast it." In every darkness, there is the potential for illumination. In every ending, a seed of beginning. The Rites in this grimoire are not the end of your journey—they are merely the next step upon a path that has no final destination, a correspondence you keep with the world's older names.

\section*{The Eternal Whisper}

As you close this book and prepare for whatever comes next, remember the fundamental truth Ikasha taught me in my first communion with her shadow:

\begin{verse}
\textit{The deepest shadows} \\
\textit{Are cast by the brightest flames.} \\
\textit{The truest vows} \\
\textit{Are written in signs and scars.} \\
\textit{The strongest Rites} \\
\textit{Are paid for twice—} \\
\textit{Once in the asking,} \\
\textit{And again when the Patron answers.}
\end{verse}

May your shadows be deep, your flames be bright, and your wisdom be greater than your power. Tend your Symbols. Mind your ledgers. Choose your Patrons as carefully as you choose your friends.

\shadow{Rukhir caws once, sharp and final, and I know that our work here is done.}

Until we meet again in the spaces between the words, dear reader, may the shadows keep you safe, and may the light never burn too brightly for your eyes.

\begin{center}
\textit{Finis}
\end{center}

\shadow{But in the world of Rites and patrons, nothing is ever truly finished. Only... paused.}

\end{document}
