\documentclass[11pt]{book}
\usepackage{geometry}
\geometry{inner=1.5cm,outer=1.5cm,top=2cm,bottom=2cm}
\usepackage{fontspec}
\setmainfont{EB Garamond}
\usepackage{titlesec}
\usepackage{tcolorbox}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xcolor}
\definecolor{wardgray}{RGB}{40,40,45}
\definecolor{candleglow}{RGB}{180,140,80}
\definecolor{shadowpurple}{RGB}{60,30,60}
\usepackage{parskip}
\usepackage{epigraph}

% Chapter style
\titleformat{\chapter}[display]
{\normalfont\Large\itshape\color{wardgray}}
{\titlerule[0.5pt]\vspace{1ex}\large\scshape\color{candleglow} Chapter \thechapter}
{0.5ex}
{\huge\bfseries\color{wardgray}}
[\vspace{0.5ex}\titlerule[0.5pt]]

\begin{document}
\input{titlepage}

\frontmatter

\thispagestyle{empty}
\vspace*{4cm}
\begin{center}
{\LARGE\itshape\color{wardgray} The Unpoured Cup:}\\[0.5cm]
{\Huge\bfseries\color{wardgray} A Guide to the Ninth and the Hollow}\\[1cm]
{\large\scshape\color{candleglow} As Told by Matron Alfiya of the Eight Doors}\\[0.5cm]
\rule{8cm}{0.5pt}\\[0.5cm]
{\small\itshape\color{wardgray} "I have kept the ninth door closed for sixty years.}\\
{\small\itshape\color{wardgray} This is what I know."}
\end{center}
\vspace*{2cm}
\clearpage

\chapter*{Preface: The House with Eight Doors}
\addcontentsline{toc}{chapter}{Preface}

It is the hour between dusk and dark, and I have lit eight candles. The ninth remains unlit. It will stay unlit until I am done speaking, and perhaps longer. The wards are tied—eight knots, each one a promise I have kept. I tell you this not to frighten you, but to prepare you.

You have come to learn about the Ninth.

The young think it is a number. The foolish think it is a superstition. The reckless think it is a source of power. They are all wrong, and they are all right, and none of them will live long enough to understand why.

I am Matron Alfiya of the Aelaerem downs. I have tended this house for sixty years. I have seen what happens when the ninth cup is poured, when the ninth step is taken, when the ninth bell rings. I have seen what the Hollow does to those who make themselves \textit{interesting}. I have kept the ninth door closed.

This book is not for the curious. It is for the \textit{wary}. It will not teach you how to use the Ninth. It will teach you how to survive it. If you read it and think you understand, you have misunderstood. The Ninth is not a puzzle to be solved. It is a boundary to be respected.

I have said my piece. The candles are burning low. Let us begin.

\begin{flushright}
— Matron Alfiya, Keeper of the Eight Doors\\
Aelaerem, in the hour before the ninth bell
\end{flushright}

\clearpage

\section*{A Note on Safety}

This guide is intended for Game Masters. It contains the deepest lore of the setting—the truths that no character should know, and that no player needs to know. Use it to inform your rulings, your atmosphere, and your consequences. \textit{Never} reveal it to players. The mystery is the point.

\tableofcontents
\clearpage

\mainmatter

\input{copyright}
\chapter{On the Number Eight}

\epigraph{Count to eight. Then stop. The ninth is not for you.}{—Aelaerem proverb}

In every culture of the Amaranthine, the pattern holds. The Aeler count their breaths in eights. The Aelinnel measure their steps in eights. The Ykrul tie their hostage strings in eights. The Linn ring their bells in eights. The Thepyrgosi climb their stairs in eights.

This is not superstition. It is not tradition. It is the oldest law written in the oldest stone, older than the Patrons, older than the empires, older than memory.

\textbf{The Rule of Eight:}

\begin{tcolorbox}[colback=gray!5!white, colframe=wardgray!75!black, title=The Rule of Eight,breakable]
\begin{itemize}
  \item \textbf{The world was made with eight seams.} The ninth seam is where it leaks.
  \item \textbf{The Patrons answer to eight names.} The ninth name is silence.
  \item \textbf{The bell tolls eight times.} The ninth bell rings in the space between heartbeats.
  \item \textbf{The cup is poured eight times.} The ninth cup is for the Hollow.
  \item \textbf{The oath is witnessed by eight.} The ninth witness is always a ghost.
  \item \textbf{The step is taken eight times.} The ninth step is never the one you planned.
\end{itemize}
\end{tcolorbox}

Why eight? Because eight is enough. Eight is the number of completion. Eight is the number of the living, the breathing, the \textit{settled}. Eight is the number of a story that has an ending.

The ninth is what comes \textit{after}. The ninth is the epilogue that refuses to end. The ninth is the sequel no one asked for.

When you take the ninth step, you step off the map. When you pour the ninth cup, you invite a guest who does not drink. When you ring the ninth bell, you ring for something that does not answer—at least, not in words you want to hear.

\subsection{Why the Number Matters}

The number eight is not magic. It is \textit{structure}. The world is held together by boundaries—between life and death, between waking and dreaming, between what is owed and what is paid. These boundaries are not walls. They are \textit{counts}. A boundary that can be counted can be crossed. A boundary that can be crossed can be exploited.

The eighth step is the final step before you reach the threshold. The ninth step is the threshold itself. When you take it, you have \textit{crossed}. The world on the other side is not the world you left. The rules are different. The debts are heavier. The witnesses are older.

The Aelaerem understand this better than most. We pour eight cups at a feast. We pour the ninth into the earth—or we do not pour it at all. The Hollow is patient. It does not need the cup. It will wait until you forget.

\chapter{On the Ninth}

\epigraph{The ninth is not a number. It is a gap.}{—Matron Alfiya}

The Ninth is the overflow condition of reality. It is what happens when the world has too much \textit{meaning}, too much \textit{intent}, too much \textit{obligation}. It is the boundary beyond which lies the hyper-real—a place where everything is true and nothing is comprehensible.

The Ninth is not sentient. It does not intend. It simply \textit{is}. Like the pressure behind a dam, it exists because the world cannot contain everything. It is the flaw in the architecture, the crack in the keystone, the extra breath that no one counted.

When a person pushes too hard—too much ambition, too many broken promises, too much \textit{attention}—they create a pressure. The Ninth is where that pressure escapes. It manifests as coincidence, as misfortune, as the sudden, unwelcome feeling that something is watching you from a place you cannot see.

\textbf{The Ninth is not evil.} It is not good. It is not a punishment. It is a \textit{consequence}. You cannot break a promise and expect the world to forget. You cannot take the ninth step and expect to stay on the same path. You cannot pour the ninth cup and expect the guest to leave.

\subsection{Manifestations of the Ninth}

The Ninth expresses itself in patterns. If you know what to look for, you can see it before it sees you.

\begin{tcolorbox}[colback=shadowpurple!5!white, colframe=shadowpurple!75!black, title=Signs of the Ninth,breakable]
\begin{itemize}
  \item \textbf{The Ninth Door.} A corridor has nine doors. The ninth leads to a room that was not on the blueprint. It is always warm. It is always waiting.
  \item \textbf{The Ninth Witness.} A count reaches nine. The ninth witness is always a ghost—or becomes one. They do not speak. They only remember.
  \item \textbf{The Ninth Cup.} A feast pours nine cups. The ninth is never drunk. It sits on the table, cooling, until the Hollow notices.
  \item \textbf{The Ninth Breath.} A person counts nine breaths aloud. The next breath is not theirs. It belongs to whoever is standing behind them.
  \item \textbf{The Ninth Name.} A name is spoken nine times. The ninth time, someone answers. It is never the person you called.
  \item \textbf{The Ninth Bell.} A bell rings nine times. The sound does not fade. It waits for an answer. It can wait forever.
  \item \textbf{The Ninth Step.} A traveler takes nine steps on a path. The ninth step is never the one they planned. It always leads somewhere unexpected.
\end{itemize}
\end{tcolorbox}

These are not omens. Omens are warnings. These are \textit{signs}. They mean the Ninth is present. They do not mean you are doomed. They mean you are \textit{noticed}.

\subsection{What the Ninth Does}

The Ninth does not harm. It \textit{complicates}. It takes a simple promise and adds a clause. It takes a straight road and adds a bend. It takes a finished story and adds one more page.

\begin{itemize}
  \item \textbf{It erodes certainty.} A character who brushes against the Ninth finds their assumptions undermined. They cannot be sure of what they remember. They cannot trust what they see.
  \item \textbf{It attracts obligation.} The Ninth is the magnet for the Hollow. A character who becomes entangled with the Ninth finds themselves \textit{owed}. Debts appear where none existed. Promises become binding. Witnesses appear to remember.
  \item \textbf{It invites interest.} The Ninth is the boundary beyond which is the hyper-real—a place where meaning is so dense it crushes. To touch the Ninth is to become \textit{interesting}. And being interesting is the most dangerous thing in the world.
\end{itemize}

\chapter{On the Hollow}

\epigraph{The Hollow is the hunger awakened by obligation.}{—Matron Alfiya}

If the Ninth is the overflow, the Hollow is what the overflow feeds. It is the accumulated weight of every promise broken, every debt unpaid, every oath sworn and forgotten. It is not a place. It is a \textit{condition}. A gravity. The urge of obligation to find its balance.

The Hollow does not pursue. It \textit{attracts}. Like calls to like. A person who carries too many broken promises finds themselves drawn to others who carry the same. A village that owes too much finds its wells dry and its children restless. A kingdom that has forgotten its treaties finds its borders shifting in the night.

The Hollow is patient. It has no reason to hurry. Every promise will be kept or broken eventually. Every debt will be paid or multiplied. The Hollow simply waits.

\subsection{The Hunger}

The Hollow is described as hungry because it is \textit{incomplete}. Every obligation creates a small emptiness in the world. The Hollow is the sum of those emptinesses. It wants to be filled. It wants to be \textit{balanced}. It will accept payment in any form—coin, blood, memory, story, or simply the acknowledgement that something is owed.

The hunger is not malicious. It is simply \textit{present}. It does not demand. It \textit{attracts}. A character who owes a debt will find themselves in situations where that debt can be paid—usually at the worst possible moment. The Hollow does not need to chase them. They will come to it eventually.

\subsection{The Hollow and the Ninth}

The Ninth is the \textit{source}. The Hollow is the \textit{current}.

When the Ninth opens, the Hollow flows through. The Ninth is the crack in the dam; the Hollow is the water that spills through. They are not the same thing, but they are inseparable. You cannot have one without the other.

The Hollow does not \textit{cause} the Ninth. The Ninth exists regardless. But the Hollow is what makes the Ninth \textit{dangerous}. Without the Hollow, the Ninth would simply be a curiosity—a strange door, an odd count, a peculiar coincidence. With the Hollow, it becomes a \textit{threat}. A claim. A summons.

\chapter{On Warding and Practice}

\epigraph{The eighth ward holds. The ninth ward is not for you to know.}{—Matron Alfiya}

The Ninth cannot be defeated. It cannot be sealed. It cannot be outrun. What it \textit{can} be is \textit{observed}. Acknowledged. Respected.

The peoples of the Amaranthine have developed a thousand small practices to keep the Ninth at arm's length. None of them are magic. They are habits. Customs. A way of reminding the world that you are \textit{settled}.

\subsection{The Practice of Eight}

\begin{tcolorbox}[colback=gray!5!white, colframe=wardgray!75!black, title=Observances of the Eight,breakable]
\begin{itemize}
  \item \textbf{Count to eight.} Then stop. The ninth is not for you.
  \item \textbf{Pour eight cups.} The ninth is for the earth, or for the Hollow. Do not drink it.
  \item \textbf{Ring the bell eight times.} The ninth bell is not for the living.
  \item \textbf{Name eight witnesses.} The ninth witness is always a ghost. Do not call them.
  \item \textbf{Walk eight steps.} Then turn, or pause, or stop. The ninth step is the threshold.
\end{itemize}
\end{tcolorbox}

These practices do not \textit{prevent} the Ninth. They \textit{negotiate} with it. They acknowledge its presence without inviting it in. They say, "I know you are there. I know what you can do. I am not ready to be interesting. I am still settling my accounts."

\subsection{The Ward of Silence}

The most powerful ward against the Ninth is silence. Not the silence of the grave—the silence of the \textit{witness}. The act of \textit{not drawing attention}.

When you break a promise, you attract the Hollow. When you tell someone about the broken promise, you attract the Ninth. The Ninth feeds on \textit{recognition}. If you keep your obligations private, you limit their reach.

This is why the Aelaerem do not speak of the Ninth. We acknowledge it. We respect it. We do not give it a name it can answer to.

\chapter{Using the Ninth and Hollow in Play}

\epigraph{The GM's job is not to punish. It is to make the world remember.}{—Matron Alfiya}

The Ninth and Hollow are not mechanics to be \textit{applied}. They are themes to be \textit{used}. They are the background texture of the world, the low hum beneath every scene. They should never be announced. They should be \textit{felt}.

\subsection{When to Invoke the Ninth}

The Ninth appears when pressure has been building. When a character has made too many promises, broken too many oaths, or simply acted with too much \textit{certainty}. The Ninth is the world's way of saying "you cannot push forever."

\begin{tcolorbox}[colback=shadowpurple!5!white, colframe=shadowpurple!75!black, title=Triggers for the Ninth,breakable]
\begin{itemize}
  \item \textbf{A pattern of broken promises.} Two or more significant oaths broken without restitution.
  \item \textbf{A pattern of ignored warnings.} Omens, superstitions, or local customs flouted.
  \item \textbf{A pattern of escalation.} Taking the ninth step, pouring the ninth cup, calling the ninth witness.
  \item \textbf{A pattern of attention.} Becoming \textit{too important}. Too interesting. Too central to the story.
\end{itemize}
\end{tcolorbox}

\subsection{Examples of the Ninth in Play}

\begin{itemize}
  \item \textbf{The Locked Room.} The players enter a room with eight doors. Each door leads to a different exit. But they count nine doors. The ninth door is not on the map. It is warm to the touch. When they open it, they are not in the same building.
  \item \textbf{The Interrogation.} The players interrogate a prisoner. They ask eight questions. The ninth question is answered by a voice that does not belong to the prisoner. The voice knows things it should not. The voice is not trying to help.
  \item \textbf{The Feast.} The players are invited to a feast. They are told to pour the ninth cup for the honored guest. When they pour it, the cup does not empty. The guest does not arrive. Something else notices the cup is waiting.
  \item \textbf{The Count.} A player counts their steps in an unfamiliar dungeon. They reach eight, then nine. The ninth step is on a different floor. There is no stair. There was never a stair. There is now.
\end{itemize}

\subsection{When to Invoke the Hollow}

The Hollow appears when an obligation has been \textit{acknowledged}. When a character knows they owe something and has not paid it. When a promise has been made but not kept.

\begin{tcolorbox}[colback=shadowpurple!5!white, colframe=shadowpurple!75!black, title=Triggers for the Hollow,breakable]
\begin{itemize}
  \item \textbf{An unpaid debt.} Coin, service, or simply attention.
  \item \textbf{An unkept promise.} An oath spoken and not fulfilled.
  \item \textbf{An unacknowledged witness.} Someone who saw something and was not thanked.
  \item \textbf{A gift given and not returned.} Hospitality is not free. It is a loan.
\end{itemize}
\end{tcolorbox}

\subsection{Examples of the Hollow in Play}

\begin{itemize}
  \item \textbf{The Debt Collector.} A character who owes a favor receives a visitor. The visitor does not ask for payment. They simply \textit{wait}. They will keep waiting until the favor is paid. They do not sleep.
  \item \textbf{The Missing Witness.} A character made a promise in private. Now they find themselves in a situation where the promise is \textit{known}. Someone else knows about it. They do not know how.
  \item \textbf{The Weight.} A character who has broken too many promises begins to feel heavy. Their pack is not heavier. Their steps are. They slow down. They tire faster. They are carrying the weight of what they owe.
  \item \textbf{The Echo.} A character who has forgotten an obligation hears a voice repeating the terms. It is their own voice. They said it once. Now it is saying it back.
\end{itemize}

\subsection{Timer Recommendations}

The Ninth and Hollow work best when they are not random. They are \textit{cumulative}. Use timers to track the pressure.

\begin{center}
\begin{longtable}{p{3cm}p{3.5cm}p{5.5cm}}
\caption{Timers for the Ninth and Hollow}
\label{tab:ninth-timers} \\
\toprule
\textbf{Timer} & \textbf{Size} & \textbf{What It Tracks} \\
\midrule
\endfirsthead
\multicolumn{3}{c}%
{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Timer} & \textbf{Size} & \textbf{What It Tracks} \\
\midrule
\endhead
\bottomrule
\endfoot
\endlastfoot
Promise Debt & 4--6 & Unkept oaths; ticks when a promise is broken or ignored. When full, the Hollow sends a reminder. \\
Ninth Attention & 6--8 & Escalation; ticks when characters take the ninth step, pour the ninth cup, or ignore warnings. When full, the Ninth manifests physically. \\
Hollow Hunger & 8--10 & Accumulated obligation; ticks when characters accept gifts, hospitality, or favors without reciprocity. When full, the Hollow collects. \\
\end{longtable}
\end{center}

\section*{A Final Warning}

I have spoken of the Ninth. I have spoken of the Hollow. I have told you how to recognize them, how to avoid them, and what to do when they find you.

But I have not told you everything. I cannot. The Ninth does not permit complete knowledge. It is the boundary beyond which is the hyper-real—a place where meaning is so dense it destroys. To know the Ninth fully is to become part of it.

This is why the Aelaerem do not write about the Ninth. We speak about it, and we stop. We pour eight cups, and we do not pour the ninth. We count eight breaths, and we do not count the ninth.

The ninth is not for you.

\begin{flushright}
— Matron Alfiya, Keeper of the Eight Doors
\end{flushright}

\appendix

\chapter*{Appendix: Quick Reference for the GM}
\addcontentsline{toc}{chapter}{Appendix: Quick Reference for the GM}

\section*{Signs of the Ninth (d8)}
\begin{enumerate}
  \item A door appears where no door was.
  \item A count reaches nine unexpectedly.
  \item A witness speaks out of turn.
  \item A sound does not fade.
  \item A reflection moves independently.
  \item A name is spoken and answered.
  \item A step lands on a different floor.
  \item A cup does not empty.
\end{enumerate}

\section*{Signs of the Hollow (d8)}
\begin{enumerate}
  \item A visitor arrives who does not leave.
  \item A promise is remembered by someone who was not there.
  \item A weight settles on the shoulders.
  \item A voice repeats what was said.
  \item A gift is offered and cannot be refused.
  \item A debt is called in with interest.
  \item A witness appears who cannot be forgotten.
  \item A silence is filled with waiting.
\end{enumerate}

\section*{GM Guidance: The Balance}

The Ninth and Hollow are not punishments. They are \textit{consequences}. They should never feel arbitrary, and they should never feel unfair. When a character attracts the Ninth or the Hollow, it should be because \textit{they did something}. They broke a promise. They ignored a warning. They made themselves interesting.

\textbf{The Golden Rule:} If a player could not have known what they were doing, do not invoke the Ninth or Hollow. Let it be a consequence of \textit{choice}, not of ignorance.

\textbf{The Second Rule:} When the Ninth or Hollow appears, it should change the situation. A complication, not an ending. The players should still have agency. They should still be able to act. They should simply have to act \textit{around} the Ninth.

\textbf{The Third Rule:} The Ninth and Hollow are not secrets to be revealed. They are textures to be felt. Do not explain them. Do not name them. Let the players discover them through experience, through whispers, through the slow accumulation of dread.

\begin{flushright}
— Matron Alfiya
\end{flushright}

\end{document}