\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{enumitem}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\begin{document}

\section*{Narethi --- The Unburied of the Deep Desert}
\textit{``The sun remembers what the sand forgets. We remember both. That is our curse. That is our strength.''}

\subsection*{Design Goals}
\begin{itemize}[leftmargin=*,nosep]
    \item \textbf{Resonance as resource:} The Unburied’s dream leaks into waking life---psionic power comes at the cost of identity erosion.
    \item \textbf{Two patrons, one people:} The \textit{Unbroken Way} teaches discipline; the \textit{Riven Path} unleashes passion. Narethi walk both, fracturing beautifully.
    \item \textbf{Thresholds of salt and bone:} Memory is currency. Ancestors vote. The Hollow watches from the dune’s edge.
    \item \textbf{Core-first:} Position/DV/Effect, timers, and Strings do the work; no new math.
\end{itemize}

\subsection*{Quickstart (2 Minutes)}
\begin{enumerate}[leftmargin=*,nosep]
    \item Choose your lens: \textit{Salt-Kin} (coastal) • \textit{Dune-Walker} (deep desert) • \textit{Sunken-Court Exile} (urban ruins) • \textit{Resonance-Touched} (psionic specialist)
    \item Mark \textbf{Resonance Leash} [6], \textbf{Memory Debt} [4], \textbf{Ancestral Favor} [4]
    \item Choose \textbf{2 Kin Gifts} (e.g., \textit{Sand-Reader, Unbound Step, Venom-Kissed, Stillness Within})
    \item Start with \textbf{1 String} (e.g., \textit{Salt-Kin token, Unburied’s whisper}) and \textbf{1 Obligation} (e.g., \textit{dream-debt, hollow-witnessed})
    \item Map \textbf{4–6 nodes}: Sunken Court • Salt Flat • Oasis of Echoes • Bone Archive • Drowned Caravanserai • The Sealed Door
    \item Open with a situation: \textit{Dream Heist} • \textit{Ancestral Vote} • \textit{Hollow Tithe} • \textit{Resonance Bleed}
\end{enumerate}

\subsection*{I. The Unburied --- Myth and Memory}
The Narethi do not bury their dead. They leave them in the salt flats or the deep caverns, where the sun and the wind and the slow, patient pressure of the earth reduce bone to memory and memory to resonance. The \textbf{Unburied}---the first of their kind, a mummified First One sealed beneath the Sunken Courts---dreams still. Its dreams leak upward.

Every Narethi is touched by this dream. It manifests as \textbf{Resonance}: a sixth sense, a psychic hum, the ability to hear what was never spoken. But the dream is hungry. It remembers. And it expects payment.

\subsection*{II. Lands After Landfall}
\paragraph{The Sunken Courts (Deep Desert)}
A subterranean city of salt, basalt, and fossilized coral. The Unburied sleeps at its heart. The Courts are ruled by \textbf{Watchers}---priest-archivists who tattoo the names of the dead onto their own skin. Outsiders are not welcome; those who enter rarely leave unchanged. \textit{Strings:} resonance-shard, sealed door token, watcher’s silence.

\paragraph{The Salt Flats (Coastal)}
Brilliant white, blinding, and treacherous. The Narethi harvest salt here, but the flats also serve as a cemetery: the dead are laid on stone tables and left to sublimate under the sun. At dusk, the flats whisper. \textit{Strings:} salt-cake, tide-echo token, mummified hand.

\paragraph{The Oasis of Echoes (Liminal)}
A freshwater spring at the edge of the deep desert. The water here carries memories---dip a cup, and you may hear a conversation from three centuries ago. The Narethi use it to consult ancestors. Non-Narethi who drink from it without offering a name are \textbf{Hollow-Marked} (the Hollow knows their face). \textit{Strings:} memory-water vial, echo-shell, ancestor’s whisper.

\paragraph{The Bone Archive (Underground)}
A library built from the remains of a thousand generations. Each bone is a scroll; each skull is a chapter. The \textbf{Bone-Scribes} maintain the Archive, adding a tooth or a knuckle for each truth recorded. The Archive is the only place where the \textbf{True Name} of the Unburied is written---but it is written in a language that no living Narethi remembers. \textit{Strings:} scribe’s needle, bone-token, sealed truth.

\subsection*{III. The Two Patrons --- Stillness and the Tide}
The Narethi do not choose one patron. They walk the \textbf{Unbroken Way} (discipline, balance, control) and the \textbf{Riven Path} (passion, fury, liberation) simultaneously. The tension between them is their theology.

\paragraph{The Unbroken Way (Stillness)}
Taught by the \textbf{Watchers}. Teaches that the mind is a fortress, that emotion is a tide to be measured, and that the Unburied’s dream must be \textit{filtered}, not embraced. \textit{Signature Technique: Resonance Cage}---once per scene, convert 2 Mental Strain into a \textbf{Stillness Token} (reduce DV of next psionic roll by 1).

\paragraph{The Riven Path (Tide)}
Taught by the \textbf{Unbound}---exiles and heretics who believe the Unburied’s dream is not a curse but an inheritance. They seek to \textit{bloom}, to let the resonance reshape them. \textit{Signature Technique: Unbound Surge}---once per scene, mark 1 Fatigue to gain +2 dice on a psionic roll; the GM gains 1 SB (the Unburied wakes a little further).

\paragraph{The Fracture} Most Narethi live between these poles. A character who favors Stillness gains +1 die to resist mental intrusion but suffers –1 die to Empathic Manipulation. A character who favors the Tide gains +1 die to Psychic Assault but marks Resonance Leash twice as fast.

\subsection*{IV. Law, Custom, and the Ancestral Vote}
\paragraph{The Cairn Council}
The dead vote. At each full moon, the Watchers assemble the \textbf{Cairn Council}: a circle of skulls that have been treated with preserving salt. The living speak the question; the dead answer by how the salt grains shift. A shift toward the speaker means yes; away means no. A spiral means the ancestors are divided, and a living mediator must be chosen.

\paragraph{Memory Debt}
Every Narethi carries a \textbf{Memory Debt} [4]. It ticks when they forget an obligation, fail to honor an ancestor, or allow a secret to be erased. When full, the Unburied sends a \textbf{Dream-Eater} (see bestiary) to collect---not a life, but a \textit{significant memory} (a name, a face, a skill point). The memory is lost permanently unless a substitute is offered (a stronger memory, a physical token, or a year of service to the Watchers).

\paragraph{The Hollow Tithe}
Once per season, the Narethi pour a cup of salt water into the desert at midnight. They speak no names. The Hollow drinks. If the tithe is forgotten, the Hollow sends an \textbf{Oath-Keeper} to the nearest settlement. The Oath-Keeper does not harm---it only \textit{reminds}. The reminder is always a death.

\subsection*{V. People \& Institutions}
\paragraph{Watchers} Priest-archivists who maintain the Bone Archive and tend the Unburied’s sleep. They are the only ones permitted to enter the deepest chamber of the Sunken Courts. A Watcher’s face is hidden behind a mask of salt-cured leather; their true name is tattooed on the inside of their own ribs.

\paragraph{Unbound (Exiles)} Narethi who broke the Memory Debt or chose the Tide over Stillness. They wander the desert, selling their psionic services to caravans and city-states. Some become assassins (the Riven Path makes them deadly); others become spies (the Unbroken Way makes them invisible). The Unbound are feared, pitied, and occasionally hired---but never welcomed home.

\paragraph{Bone-Scribes} Archivists of the Cairn Council. They do not speak aloud; they write on bone with needles dipped in their own blood. A Bone-Scribe can recall any Memory Debt recorded in the Archive---for a price (a tooth, a lock of hair, a promise to forget something else).

\paragraph{Dream-Eaters (Spirits)} Manifestations of the Unburied’s hunger. They appear as jackals with too many eyes, or as reflections that move independently. They do not kill; they \textit{collect}. When a Dream-Eater takes a memory, the victim forgets not the fact but the \textit{feeling}. They remember that they loved someone, but not the warmth of that love. They remember that they were betrayed, but not the sting. This is the Hollow’s true cruelty: not death, but emptiness.

\subsection*{VI. Aesthetics \& Speech}
\paragraph{Colors} Salt-white, basalt-black, dried-blood red, the violet of the steppe’s edge.
\paragraph{Shapes} Spirals, mazes, the eight-pointed star of the Unbroken Way (the ninth point is the Unburied’s eye).
\paragraph{Clothing} Light robes woven from desert flax, bone needles, salt-encrusted veils. Watchers wear masks; Unbound wear none.
\paragraph{Proverbs}
\begin{itemize}[leftmargin=*,nosep]
    \item “The sun remembers. The sand forgets. We remember both.”
    \item “A memory unpaid is a door left open.”
    \item “Stillness is not silence---it is the breath before the strike.”
    \item “The Riven Path does not forgive; it burns clean.”
    \item “When the ancestors vote, listen. They have all the time in the world.”
\end{itemize}

\subsection*{VII. Rivals \& Kindreds}
\paragraph{The Aeler (Dwarves)} Respect the Narethi’s breath-counting, distrust their psionics. Aeler holds employ Unbound as “resonance auditors” but never let them near the keystone. \textit{Relationship:} tense trade; shared hatred of the Hollow.

\paragraph{The Ykrul} The steppe clans call the Narethi “Salt-Eaters” and avoid the deep desert. A Ykrul who accepts a Narethi’s guest-cup is considered \textit{resonance-marked}---the Hollow may visit their dreams. \textit{Relationship:} wary respect; occasional alliances against common foes.

\paragraph{The Sidhi (Brass Coast)} Kindred spirits. Both peoples carry the weight of a forgotten empire. Narethi and Sidhi trade in memories, salt, and lighthouse charts. Some Sidhi lighthouses are built on Narethi salt-pans; the arrangement is old and rarely spoken of. \textit{Relationship:} old compact, weathering.

\subsection*{VIII. Festivals \& Omens}
\paragraph{The Unburied’s Wake (annual)} A night of silence. No fires. No speech. The Narethi sit in the dark and \textit{remember}---each person reciting the names of their ancestors in their head. At dawn, they pour salt water on the ground. If the water steams, the Unburied has accepted the tribute. If it remains cold, a Dream-Eater is already hunting.

\paragraph{The Sand Recital (seasonal)} Bone-Scribes gather to recite the Memory Debts of the previous season. They do not speak; they tap the rhythm on hollow bones. The audience listens with their hands pressed to the ground, feeling the vibration. A missed beat means a Debt is overdue. The Hollow will send a reminder within a week.

\paragraph{Omens that matter}
\begin{itemize}[leftmargin=*,nosep]
    \item A well that tastes of iron---the Unburied is dreaming of thirst.
    \item A reflection that moves after you stop---a Dream-Eater is considering you.
    \item A bone whistle found in the desert---someone has paid the Hollow with a name.
    \item Salt crystals that form in the shape of a spiral---the ancestors are voting.
\end{itemize}

\subsection*{IX. Adventure Seeds}
\begin{enumerate}[leftmargin=*,nosep]
    \item \textbf{The Dream Heist.} The Unburied’s dream has shifted. A fragment of its memory---the location of a pre-Silence weapon---is leaking into the resonance. The Watchers want it sealed. The Unbound want to sell it. The party must enter the dream (Obishaal) and retrieve the memory before the Hollow does.
    \item \textbf{The Ancestral Vote.} A Cairn Council is deadlocked (spiral). The party must act as living mediators---investigating an old murder, a broken oath, or a forgotten debt---and present evidence to the skulls. The dead are patient. They will wait until the living get it right.
    \item \textbf{The Salt-Thief.} Someone is harvesting salt from the flats without offering names to the Hollow. The Hollow is angry. The party must find the thief and ensure the names are paid---or offer names of their own. The thief may be desperate, or may be a rival faction trying to starve the Narethi of memory.
    \item \textbf{The Unbound Uprising.} A faction of exiles has returned to the Sunken Courts, demanding that the Unburied be awakened---not to destroy, but to \textit{speak}. The Watchers have sealed the deepest chamber. The party must choose: help the Unbound crack the seal, or help the Watchers drive them back into the desert.
    \item \textbf{The Hollow’s Tithe.} The village forgot the seasonal offering. An Oath-Keeper has appeared at the well. It does not speak; it only waits. The party must pay the tithe---with a memory, a name, or a story---before the Oath-Keeper takes one of the villagers. But the Oath-Keeper is patient. It will accept a substitute, but the substitute must be \textit{meaningful}.
\end{enumerate}

\subsection*{X. The Resonance Leash (Core Mechanic)}
Every Narethi character tracks a \textbf{Resonance Leash} [6]. It advances when:
\begin{itemize}[leftmargin=*,nosep]
    \item You use a significant psionic ability (Telepathy to read deep memories, Clairvoyance to scry, Psychic Assault, etc.) – mark +1.
    \item You fail a Resolve test while in a thin place or near a Dream-Eater – mark +1.
    \item The GM spends 2+ Story Beats to represent the Unburied’s dream shifting – mark +1.
\end{itemize}
When the Leash fills, you experience a \textbf{waking vision} of the First Scouring---a fragment of apocalyptic memory (a city of black glass, a sky on fire, a scream without end). You mark \textbf{1 Fatigue} (the strain of the memory), and the Leash resets to 0.

\paragraph{Vision Consequences (d6)}
\begin{enumerate}[leftmargin=*,nosep]
    \item You glimpse a future death (your own or an ally’s). Gain a \textbf{Fate Token} (once per session, reroll a failed roll against that death).
    \item You hear a true name---of a creature, a place, or a patron. The GM must whisper it to you. You may use it once as a String.
    \item Your shadow separates from you for one scene. It acts as a \textbf{Cap 1 follower} (can scout, distract, or steal small objects). It returns when you sleep.
    \item You forget a minor memory (the face of a passerby, a childhood pet’s name). The GM gains 1 SB to spend on a memory-related complication.
    \item You speak a prophecy aloud (GM gives you one sentence about the future). It is always true, but cryptic. Mark 1 Obligation to the Unburied.
    \item A Dream-Eater appears at the edge of the scene. It does not attack---it watches. If you offer it a memory (erase a minor fact), it leaves. If you ignore it, it follows you for the session.
\end{enumerate}

\subsection*{XI. Kin Gifts (Pick 2)}
\paragraph{Sand-Reader (Minor, 2 XP)} You can read the memory of a place by tasting a pinch of its sand or soil. Once per session, ask the GM one yes/no question about an event that occurred within the last week within Close range.

\paragraph{Unbound Step (Minor, 2 XP)} You ignore the first level of difficult terrain caused by sand, salt, or loose stone. Once per scene, you may treat a fall as if you had taken no damage (landing in a slide).

\paragraph{Venom-Kissed (Minor, 2 XP)} You are immune to mundane poisons and gain +1 die to resist magical poisons. Your own saliva is mildly toxic; you may deliver a Poisoned Condition (DV 3) via a bite or kiss.

\paragraph{Stillness Within (Minor, 2 XP)} Once per scene, as a reaction, you may reduce incoming Mental Strain from a psionic ability by 1 (minimum 0). You cannot use this if your Resonance Leash is already at 5+.

\paragraph{Memory Eater (Major, 6 XP)} When you touch a willing or helpless creature, you may drain one specific minor memory from them (the name of a contact, a password, a childhood fear). You gain 1 Boon; the target forgets that memory permanently. Using this on an unwilling target requires a successful Psychic Assault roll.

\paragraph{Cairn Speaker (Major, 6 XP)} You may commune with the dead through a physical anchor (a skull, a bone, a lock of hair). Once per session, ask the GM one question about the dead person’s life, death, or unfinished business. The GM must answer truthfully, though cryptically. Mark 1 Fatigue after use.

\paragraph{Resonance Cage (Prestige, 10 XP)} You may suppress your own psionic signature completely. For one scene, you cannot be detected by any psionic means (Clairvoyance, Telepathy, Empathic Manipulation). You also cannot \textit{use} psionics during that scene. Once per session.

\paragraph{Unbound Surge (Prestige, 10 XP)} When your Resonance Leash fills, you may choose to \textit{not} reset it. Instead, you gain +2 dice to all psionic rolls for the next scene. At the end of the scene, the Leash fills again---this time, you experience a \textit{major} vision (roll twice on the Vision Consequences table and take both). You cannot use this talent again until you have cleared a Memory Debt or performed a Hollow Tithe.

\subsection*{XII. Strings \& Secrets}
\paragraph{Strings (Sample)}
\begin{itemize}[leftmargin=*,nosep]
    \item \textbf{Salt-Kin Token} – Proof of shared ancestry; once per session, gain +1 Position in a Narethi settlement.
    \item \textbf{Ancestor’s Whisper} – A bone-whistle that, when blown, calls a minor ancestor spirit to answer one question (the spirit is cryptic and demands a memory in return).
    \item \textbf{Watcher’s Seal} – A leaden seal stamped with a spiral; shows you are under the Watchers’ protection (or scrutiny).
    \item \textbf{Dream-Eater’s Tooth} – A small, warm tooth. Once, you may offer it to a Dream-Eater in place of a memory. The Dream-Eater will leave---but the tooth returns to your pocket the next dawn.
\end{itemize}

\paragraph{Secrets (Sample)}
\begin{itemize}[leftmargin=*,nosep]
    \item The Unburied is not a First One---it is a \textit{vessel}. The First One died during the Great Silence. The Unburied is a substitute, stitched together from the memories of a thousand dead Narethi.
    \item The Hollow’s true name is written in the Bone Archive---but reading it would wake the Unburied permanently.
    \item The Sunken Courts are not a ruin. They are a \textit{ship}. The Narethi did not flee across the Great Wake; they fell. The Courts are the wreck.
\end{itemize}

\subsection*{XIII. Running Narethi PCs}
\paragraph{Resonance as Resource} Treat the Leash as a pressure gauge, not a punishment. When it fills, offer the player a choice: take the vision (Fatigue and a complication) or \textit{bloom} (gain a temporary boon but mark a permanent Obligation). Blooming is how Narethi become Unbound.

\paragraph{Memory as Currency} Use Memory Debt to drive stories. A Debt that fills should not be a random penalty---it should be a \textit{hook}. The Hollow takes something the player actually cares about (a name, a relationship, a key piece of information). Then the player has to decide: let it go, or go after it.

\paragraph{The Two Patrons in Play} Let the player choose which patron’s voice is louder in a given scene. A Watcher might offer a \textbf{Stillness Token} (reduce Mental Strain). An Unbound exile might offer a \textbf{Tide Token} (bonus dice at the cost of a Leash tick). The tension between discipline and passion is the Narethi’s core drama.

\subsection*{XIV. Narethi in the Wider World}
\paragraph{In Ecktoria} Narethi are “dream-readers,” employed by the Censorate to verify testimony (does the witness believe they are telling the truth?). They are distrusted but necessary. A Narethi who claims to speak for the dead is ignored; one who offers to audit a memory is paid.

\paragraph{In Thepyrgos} The Synod licenses Narethi psions as “resonance consultants.” The Chain-Lanterns watch them closely. A Narethi who oversteps (e.g., reading a clerk’s mind without permission) is subject to the same candle-smoke test as any other unlicensed mage. Many Narethi in Thepyrgos work as archivists or witnesses, not as magi.

\paragraph{In the Black Banners} Narethi psions are prized for their ability to interrogate prisoners without leaving marks. They are also feared---a Narethi who learns a captain’s secrets is a Narethi who may sell them to the other side. The Banner’s solution: pay them well, keep them fed, and never let them meet the quartermaster alone.

\paragraph{In the Valewood} The Lethai-al are wary of Narethi. Both peoples carry the memory of the Great Silence, but the Lethai blame the Narethi for it (or vice versa; the record is unclear). A Narethi who enters the Valewood is offered guest-right---coldly, precisely, and for exactly three nights. No more.

\subsection*{XV. Sample Narethi Characters}
\paragraph{Eight-of-Salt (Unbound Exile)}
Concept: A Narethi who let her Resonance Leash fill one too many times. She remembers seven of the eight visions. The eighth is the one the Hollow took. She is now an exile, selling her psionic talents to caravans. \textit{Kin Gifts:} Unbound Step, Resonance Cage. \textit{Bond:} Owes a memory to a Dream-Eater that spared her. \textit{Complication:} The Hollow knows her true name.

\paragraph{The Still Scribe (Watcher Initiate)}
Concept: A Bone-Scribe who has never left the Sunken Courts---until now. She carries the Archive’s memory in her bones. She is searching for a lost truth (a clan’s name, a burial site, a forgotten sin). \textit{Kin Gifts:} Stillness Within, Cairn Speaker. \textit{Bond:} Watcher’s Seal (Protection/Obligation). \textit{Complication:} Cannot speak aloud in the presence of the dead (her voice disturbs their vote).

\paragraph{The Hollow-Touched (Accidental Psion)}
Concept: A merchant, a guide, a thief---not a trained Narethi, but exposed to the Unburied’s dream through a thin place. Now she hears the dead whispering, and the Hollow knows her face. She is not a believer. She is a survivor. \textit{Kin Gifts:} Memory Eater, Sand-Reader. \textit{Bond:} Owes a life-debt to an Unbound exile who saved her from a Dream-Eater. \textit{Complication:} Every night, she dreams of the First Scouring. She wakes with salt on her lips.

\end{document}
