\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{fontawesome5}

% Page setup
\geometry{margin=1in}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5em}

% Section formatting
\titleformat{\section}{\large\bfseries\uppercase}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\itshape}{}{0em}{}

% List formatting
\setlist[itemize]{left=0pt .. 1.5em, itemsep=0.25em}
\setlist[enumerate]{left=0pt .. 1.5em, itemsep=0.25em}

% Custom commands
\newcommand{\sectionrule}{\vspace{0.5em}\noindent\textcolor{gray}{\rule{\textwidth}{0.4pt}}\par\vspace{0.5em}}
\newcommand{\subsectionrule}{\vspace{0.3em}\noindent\textcolor{gray}{\rule{\textwidth}{0.3pt}}\par\vspace{0.3em}}
\newcommand{\checkmark}{$\checkmark$}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Fantasy City Expansion: Violets \& Stone},
    pdfauthor={Fate's Edge Team},
}

\urlstyle{same}

\begin{document}

\begin{center}
    {\LARGE \textbf{Fantasy City Expansion: Violets \& Stone}} \\
    {\large (Fate's Edge v0.4)} \\
    \vspace{0.5em}
    A plug-and-play urban module for Fate's Edge campaigns set among bridges, aqueducts, rites, caravans, and crown politics.
\end{center}

\sectionrule

\section*{Design Goals}
\begin{itemize}
    \item City as character. Districts and institutions carry tags, strings, and clocks that push back.
    \item Plug into core. Use Position/DV, SB, and clocks; no new dice math.
    \item Street to state. Blend casework, capers, and civic settlements.
    \item Low bookkeeping. Track names and tags; let clocks and strings carry weight.
\end{itemize}

\sectionrule

\section*{Quickstart (2 minutes)}
\begin{enumerate}
    \item Pick a City Kit (\S7). Set Mandate 2 / Crisis 2 unless fiction says otherwise.
    \item Choose 4 Factions from the kit; copy their Strings; start Influence [6] / Stability [6--8] / Exposure [6].
    \item Mark 2 volatile stakes on the relationship map.
    \item Draw the Urban Deck (\S6.2): one Incident (Spade), one Person (Heart), one Complication (Club) to spark the opening dilemma.
    \item Frame the first Situation Clock [4--8] and at least one Counter-Clock [4--6].
    \item Play scenes. All 1s generate SB; spend city-flavored Social SB (\S3.4).
\end{enumerate}

\textbf{Pro Tip:} Start with the Three-Clock Rule—only track the Situation Clock, one Counter-Clock, and one Campaign Clock at a time.

\sectionrule

\section*{1) City Sheet (Template)}
\textbf{[CITY NAME]} \\
Mandate (0--6): public legitimacy of the company in this city \\
Crisis (0--6): ambient pressure (riots, audits, plague, war drums) \\
Districts (5--8): each with tags \& a District Effect \\
Institutions (3--6): councils, courts, guilds, rites; each has Strings \\
Factions (4--8): playset; use Influence / Stability / Exposure clocks \\
Relationship Map: arrows with stakes; mark two volatile \\
Rites \& Laws: 1--2 citywide oaths, bans, or permissions that change Position \\
Chase Modes: foot • canal • rooftops • alleyways (DV ladders in \S5) \\
Events: festival/omen/weather tables \\

\textbf{District Effect (per scene):} Acting with a district's tags → Position +1 or DV --1 once/scene. Acting against a district's tags → Position --1 or GM banks +1 SB.

\sectionrule

\section*{2) District Tags \& Effects (pick 6--8 common)}
\begin{itemize}
    \item \textbf{Harbor \& Quays} — brackish, fog, bribes, cranes \\
    Effect: +1 Position for Smuggle/Survey; --1 Position for Sanctify/Expose. \\
    Hazard: Night-lantern failures; docksiders rumor clock.
    
    \item \textbf{Market Ward} — hucksters, scales, guild seals \\
    Effect: DV --1 for Broker/Petition with writs; +1 SB on Infiltrate (watchful). \\
    Hazard: Guild watch can ``price'' entry (concession).
    
    \item \textbf{Temple Quarter} — bells, incense, oaths \\
    Effect: Shield/Host Rite start Controlled; Bribe starts Desperate. \\
    Hazard: Breaking hospitality ticks Exposure +1.
    
    \item \textbf{Scholars' Quarter} — scriptoriums, apprentices, disputes \\
    Effect: Research/Expose DV --1 with proof; Blackmail Position --1 (they lawyer up). \\
    Hazard: Scholastic feud counter-clock.
    
    \item \textbf{Underways \& Sewers} — echo, mildew, sigils, smugglers \\
    Effect: Infiltrate DV --1; on any 1, start Sewer Folk or Wards Stir clock. \\
    Hazard: Flood surge; ward backlash.
    
    \item \textbf{Legion/Watch Barracks} — drill, pennons, billets \\
    Effect: Petition (order) starts Controlled; Expose starts Desperate. \\
    Hazard: Curfew lowers Position by one step.
    
    \item \textbf{Bridges \& Canals} — mossy stone, toll bells \\
    Effect: Chase (boat/rooftop) DV --1; Broker worsens if tolls unpaid. \\
    Hazard: Bridge-lords demand dues (Favor spend or stall).
    
    \item \textbf{Patrician Heights} — courtyards, old blood \\
    Effect: Petition DV --1 with Standing; otherwise Blackmail gains +1 SB. \\
    Hazard: Duels and scandal escalate quickly.
    
    \item \textbf{Foundries \& Yards} — smoke, sparks, strikes \\
    Effect: Intimidate/Strike-bust Position +1; Host Feast Position --1. \\
    Hazard: Rolling strike counter-clock.
    
    \item \textbf{Bazaars \& Caravanserais} — spices, dialects, caravan law \\
    Effect: Broker (trade) DV --1; Expose Position --1 (outsider loyalties). \\
    Hazard: Caravan feud ignites on SB spends.
    
    \item \textbf{Docks \& Wharves} — salt spray, rigging, bilge \\
    Effect: +1 Position for Smuggle/Infiltrate via water; --1 Position for Petition (land-based). \\
    Hazard: Storm surge; ship fires.
    
    \item \textbf{Carnival Quarter} — masks, games, pickpockets \\
    Effect: +1 SB on Carouse for Rumors; Disguise starts Controlled. \\
    Hazard: Festival crush; pickpocket networks.
\end{itemize}

\sectionrule

\section*{3) City Procedures}
\subsection*{3.1 Scene Framing}
State the question; pick a Situation Clock; set Position \& DV from fiction + district effect; start a Counter-Clock if opposition is active; tag the venue (Audience/Institution).

\subsection*{3.2 Strings (Urban)}
Common Strings: Bridge Tolls, Aqueduct Permits, Ferry Passbooks, Oath-Tablet Registry, Harbor Pilots, Festival Licenses, Censor's Seal, Ship Papers, Privateer Charters, Contraband Manifests. \\

Acting with/through a relevant String: Position +1 or DV --1 once/scene. \\
Cut/Seize a String with a 4--6 clock; on fill, move the String or strike it. \\

\textbf{Cutting Strings Example:} The characters discover forged ship papers used by smugglers. They can spend a scene cutting this string (create a [4] clock). On completion, they either seize the papers for their own use or destroy them to prevent others from using them.

\subsection*{3.3 City Dials}
\begin{itemize}
    \item \textbf{Mandate 0--6 (public legitimacy):} at 4--5, cancel one Social SB in public venues once/session; public offices DV --1. At 6, start public scenes Controlled unless fiction forbids.
    \item \textbf{Crisis 0--6 (pressure/backlash):} at 4--5, first failed public scene ticks party Exposure +1; at 6, start with a crowd Counter-Clock [4].
\end{itemize}

\textbf{New:} Mandate/Crisis Shifts. At the end of each session, if the total SB spent exceeds 10, increase Crisis by 1. If the party successfully resolved 2+ major clocks, increase Mandate by 1.

\subsection*{String Cutting Quick Reference:}
\begin{itemize}
    \item \textbf{Cost:} Dedicate one scene action to cutting/seizing
    \item \textbf{Clock:} Create 4--6 segment clock based on string importance
    \item \textbf{Resolution Options:}
    \begin{itemize}
        \item Seize: Gain control of string for your use
        \item Destroy: Prevent others from using it
        \item Redirect: Change who controls it
    \end{itemize}
    \item \textbf{Example:} Cutting "Harbor Warrants" [4] - on completion, either gain legitimate papers or prevent harbor from issuing warrants against you
\end{itemize}

\subsection*{3.4 Social SB Menu (City-flavored)}
Faux pas title/ritual; lose 1 Favor in venue • leverage leak → target/party Exposure +1 • crowd turns (Warm→Skeptical→Hostile) • permit pulled (lose a String access this act) • patrol arrives biased • festival crush • toll chain drops • ship quarantine • cargo inspection • guild audit • noble scandal • temple censure • watch curfew • harbor closure • price fixing • counterfeit goods • debt collector • bounty hunter • informer betrayal • rival sabotage.

\subsection*{Investigation Clock Management:}
\begin{itemize}
    \item Start at [6] for major mysteries
    \item Advance by 1 for Clue Grades D/C
    \item Advance by 2 for Clue Grades B/A  
    \item Advance by 3 for Clue Grades A+
    \item Truth Revelation: When filled, describe the complete truth but introduce a new complication
    \item Example Progression: Clue D → Clue C → Clue B → Clue A+ = 1+1+2+3 = 7 segments (truth revealed)
\end{itemize}

\subsection*{3.5 Between-Session Cycle}
Update world state → draw events → resolve off-screen orders → recompute dials → prep next Situation (see \S6 for events and omens).

\sectionrule

\section*{4) Urban Actions (additions)}
\begin{itemize}
    \item \textbf{Invoke Rite (Wits/Presence + Rite):} Claim hospitality, sanctuary, or festival privilege. Hit: protective tag; miss: Exposure +1.
    \item \textbf{Grease Palms (Wits + Streetwise):} Secure permits/tolls. Hit: tag Greased Palms; miss: rumor of bribery (Exposure +1).
    \item \textbf{Audit Ledger (Wits + Investigation):} Catch graft/tithe anomalies. Hit: DV --1 on later Expose.
    \item \textbf{Call the Watch (Presence + Command):} Leverage order. Hit: Watchful Patrols tag; miss: they arrive biased.
    \item \textbf{Smuggle (Wits + Subterfuge):} Move goods/people across wards. Hit: gain Leverage (1) or fill; 1s start Contraband Trail.
    \item \textbf{Carouse for Rumors (Presence + Performance):} Trade coin and stories. Hit: bank 1 Clue (\S6.1); miss: SB → Hangover Debts.
    \item \textbf{Forge Documents (Wits + Tinker):} Create false papers/permits. Hit: tag Forged Papers; miss: Exposure +1 (counterfeit discovered).
    \item \textbf{Navigate Politics (Presence + Command):} Maneuver through civic structures. Hit: DV --1 on next Petition; miss: political rival gains Influence.
    \item \textbf{Bribe Official (Wits + Sway):} Pay for favors/permissions. Hit: tag Bribe Accepted; miss: Exposure +1 (bribe rejected/extorted).
    \item \textbf{Commission Contract (Presence + Broker):} Hire services/mercenaries. Hit: tag Contract Secured; miss: Exposure +1 (contract dispute).
\end{itemize}

\sectionrule

\section*{5) Chases in Tight Streets (DV Ladders)}
Pick a mode; on a hit, advance Pursuit (or Escape) clock by effect.
\begin{itemize}
    \item \textbf{Foot:} DV 2 market • DV 3 alleys • DV 4 rooftops • DV 5 through procession.
    \item \textbf{Boat/Canal:} DV 2 open run • DV 3 under bridges • DV 4 against tide • DV 5 in fog storm.
    \item \textbf{Mount/Cart:} DV 2 quay road • DV 3 wagon traffic • DV 4 stair-streets • DV 5 temple steps.
    \item \textbf{Ship-to-Ship:} DV 2 calm waters • DV 3 choppy seas • DV 4 storm winds • DV 5 naval battle.
\end{itemize}
\textbf{Complications on 1s:} overturned barrow, toll chain drops, lantern goes out, festival throng, rigging tangles, hull breach, powder keg, enemy boarding party, reef strike, sudden fog, cargo shift, crew mutiny.

\sectionrule

\section*{6) Urban Generators}
\subsection*{6.1 Clues \& Casework}
Grades: A+ (witnessed oath; sealed writ) • A (signed testimony; verified mark) • B (two-source rumor; ledger trail) • C (hearsay; graffiti) • D (single rumor) • F (planted tale; forged writ). \\
A day without follow-up degrades one step if contested. \\
Spend: 1 Clue → DV --1 on a related Expose/Petition once. \\

\textbf{Investigation Clock [6]:} Track long-term investigations. Each significant clue advances by 1. When filled, the truth is revealed but attracts attention.

\subsection*{6.2 Urban Deck (52-card)}
Draw: Spade (Incident), Heart (Person), Club (Complication), Diamond (Boons/Resolution).
\begin{itemize}
    \item \textbf{Spades:} ritual misfire • theft of a writ • bridge collapse • missing tax barge • assassin in procession • cursed play • guild strike • seditious broadsheet • shrine desecration • plague rumor • treason trial • flood omen • Ace: cover for a coup.
    \item \textbf{Hearts:} harbormaster • guild factor • inquisitor • bridge-lord • mummer • tribune • ferry pilot • matron • archivist • street saint • exile heir • quartermaster • Ace: a most trusted patron is complicit.
    \item \textbf{Clubs:} curfew • festival crush • foreign immunities • taboo ground • closed shop • informant vanishes • watch obstruction • blackmail • riot brewing • double agent • cursed weather • Ace: solving it ruins someone you love.
    \item \textbf{Diamonds:} sealed archive • tariff exemption • festival license • street mandate • protection writ • house favor • council seat • tithe forgiving • pardon • endowment • patrol command • treasury award • Ace: The Truth, at public cost.
\end{itemize}
Clock Size by Highest Rank: 2--5 → [4], 6--10 → [6], J/Q/K → [8], A → [10].

\subsection*{6.3 Events, Omens, Weather}
Festivals: Boat Blessing • New Charter Day • Founders' Procession • Lantern Vigil • Moon Market • Salt Blessing • Ship Launch • Merchant's Fair • Sailor's Wake • Governor's Ball • Execution Day • Tax Collection. \\

Omens: blood tide • comet • cracked idol • whisper-plague • augur's scandal • ship's bell tolling • raven's call • broken mast • cursed cargo • ghost ship sighted • temple weeping • market crash. \\

Weather: fog bank • heat wave • black rain • river in spate • winter of knives • sudden storm • red sky • calm seas • waterspout • ice floes • sandstorm • hurricane winds.

\sectionrule

\section*{7) City Kits (Playsets)}
\subsection*{Faction Relationship Matrix (Zakov Example):}
\begin{itemize}
    \item Ironmasters vs Corsair Republic: Competitive (territory/control)
    \item Harbor Court vs Salt Guild: Cooperative-Adversarial (trade dependency with price disputes)
    \item All factions vs Independents: Suspicious but opportunistic
    \item Player Alignment Tips:
    \begin{itemize}
        \item 1--2 shared aims = Favorable Position
        \item 3+ shared aims = Alliance (Influence gain on success)
        \item Opposed aims = Hostile Position (--1 die or +1 SB)
    \end{itemize}
\end{itemize}

\subsection*{7.1 Mid Ahkaz — Ykrul Trade Capital of the Violet Steppe/Meadows}
Vibe: violet grasses in wind-bands, caravan gold, contract law, oasis politics, foreign tongues. \\
Districts: Violet Bazaar; Steppe Gate; Satrap Gardens; Coin-Weigh House; Under-Arcades; Caravan Rings. \\
Institutions: Ykrul Coin-Weigh (tariff court); Oasis Compact (water rights council); Steppe Rangers (patrol and feuds). \\
Factions:
\begin{itemize}
    \item \textbf{Coin-Weigh Tribunal (Tier crown):} Aim—stabilize tariffs; Strings—Tariff Seals, Coin Scales, Audit Teams.
    \item \textbf{Oasis Clans Compact (Tier city):} Aim—protect wells; Strings—Water Shares, Caravan Well Rights, Desert Guides.
    \item \textbf{Foreign Factors' League (Tier guild):} Aim—preferential duties; Strings—Warehousing, Bonded Guards, Letters of Credit.
    \item \textbf{Sand-Road Brotherhood (Tier street→guild):} Aim—own the under-arcades; Strings—Contraband Vaults, Night Porters, False Seals.
\end{itemize}
Special Rule — Steppe Mandates: In Bazaars/Caravan Rings, Broker DV --1. Exposing fraud without a Coin-Weigh Seal starts Exposure +1. \\
Adventure Starters: Missing Water Share deeds flood the market • Steppe Rangers feud spills into the Violet Bazaar • A cursed dye caravan seeks sanctuary under the Satrap Gardens.

\subsection*{7.2 Ecktoria — Marble \& Fire (Post-Imperial Capital)}
Vibe: fallen grandeur, aqueducts, legions remnant, exarch audits. \\
Districts: Grand Forum; Aqueduct Heights; Legion Quarter; Foundries; Riverside Tenements; Consecrated Courts. \\
Institutions: Imperial Exarchate (audit/oversight); Merchant Courts; Sun-Temples; Civic Aqueducts Office. \\
Factions:
\begin{itemize}
    \item \textbf{Exarchate (Tier crown):} Aim—rebalance old grants; Strings—Audit Writs, Tax Ledgers, Censures.
    \item \textbf{Legions Remnant (Tier city):} Aim—restore stipends; Strings—Veteran Rolls, Armory Keys, Drill Grounds.
    \item \textbf{Merchant Courts (Tier guild):} Aim—secure monopolies; Strings—Guild Charters, Bridge Tolls, Import Licenses.
    \item \textbf{Sun-Temple Synod (Tier synod):} Aim—purify rites; Strings—Sanctuary, Procession Routes, Oath Tablets.
\end{itemize}
Special Rule — Aqueduct Integrity: Any Expose of water theft is public: on hit, Mandate +1; on miss, Exposure +1 (panic). \\
Adventure Starters: A lost aqueduct patent resurfaces • Legions demand arrears from the treasury • A relic procession goes off-script into riot.

\subsection*{7.3 Silkstrand — City of Bridges}
Vibe: fog, dyes, masks, mummers; everything happens on a bridge. \\
Districts: Bridges \& Canals; Dye Yards; Archivolt; Playhouse Row; Pilots' Quays; Sestiere Courts. \\
Institutions: Dye Syndicate; Archivolt Curia; Bridge-Lords Council; Riverine Patrol; Playhouse Guild. \\
Factions:
\begin{itemize}
    \item \textbf{Dye Syndicate (Tier guild):} Aim—maintain pigment monopoly; Strings—Vat Houses, Guild Scales, Apprentice Oaths.
    \item \textbf{Bridge-Lords (Tier city):} Aim—rule crossings; Strings—Toll Chains, Bridge Charters, Night Keys.
    \item \textbf{Riverine Patrol (Tier city):} Aim—curb smuggling; Strings—Patrol Boats, Fog Bells, Harbor Warrants.
    \item \textbf{Playhouse Mummers (Tier street→guild):} Aim—public favor; Strings—Masque Licenses, Satire Broadsheets, Crowd Hooks.
\end{itemize}
Special Rule — Fog Bells: In fog, Chase (boat/rooftop) DV --1 but Expose Position --1; a hit satire can create Audience tags citywide. \\
Adventure Starters: A satire names names • The Night Keys list leaks • A foreign pigment line undercuts the vats.

\subsection*{7.4 Thepyrgos — Synod \& Collegium}
Vibe: law and rites braided; harbor guilds and a scholastic war. \\
Districts: Synod Hill; Collegium of Keys; Harbor Quarter; Law Courts; Pilgrims' Steps; Scriptorium Lanes. \\
Institutions: Archons' Synod; Collegium of Keys; Harbor Guild; Ember-Temple. \\
Factions:
\begin{itemize}
    \item \textbf{Archons' Synod (Tier crown):} Aim—codify chapterhouse votes; Strings—Writs of Law, Censure, Clerks.
    \item \textbf{Collegium (Tier synod):} Aim—protect privileges; Strings—Oath Registry, Sealed Stacks, Licentiae.
    \item \textbf{Harbor Guild (Tier guild):} Aim—tariffs/pilots; Strings—Pilot Licenses, Quay Masters, Lantern-Law.
    \item \textbf{Ember-Temple (Tier synod):} Aim—zeal/charity; Strings—Processions, Sanctuary, Inquisitors.
\end{itemize}
Special Rule — Ritual Hospitality: Begin with rites: Broker starts Controlled; breaking rites adds Exposure +1. \\
Adventure Starters: Charter revision threatens the Collegium • Temple schism spills into streets • A curse knits into tithe ledgers.

\subsection*{7.5 Zakov — Iron Haven (Freeport of the Crimson Straits)}
Vibe: salt-spray democracy, corsair republics, ship-forged law, maritime merchant princes. \\
Districts: Ironworks; Salt Quarter; Smugglers' Cut; Old Citadel; Copper Market; Blackwater Quays. \\

\textbf{Lore:} Zakov sits at the mouth of the Iron River where it meets the Crimson Straits, a natural harbor that has become the most lawless port in the known world. What began as a trading post for river merchants became a haven for privateers, pirates, and those fleeing justice. The city operates under the "Code of the Iron Trident"—a maritime law system where ship captains hold sovereign authority over their vessels and the harbor waters. \\

The city is divided into "Holds"—territories claimed by different corsair fleets, merchant houses, and criminal syndicates. The only unifying force is the Harbor Master's Court, which maintains the harbor's navigational aids and collects the "Iron Tithe" from all ships that dock. \\

Institutions: Ironmasters' Collegium; Corsair Council; Harbor Master's Court; Salt Guild. \\
Factions:
\begin{itemize}
    \item \textbf{Ironmasters (Tier guild):} Aim—secure raw ore/export tithes; Strings—Foundry Licenses, Ore Contracts, Bonded Guards.
    \item \textbf{Corsair Republic (Tier street→guild):} Aim—maintain maritime law/territory; Strings—Ship Papers, Privateer Charters, Harbor Rights.
    \item \textbf{Harbor Master's Court (Tier city):} Aim—keep harbor navigable/profitable; Strings—Navigation Marks, Docking Rights, Iron Tithe.
    \item \textbf{Salt Guild (Tier guild):} Aim—control salt trade/preservation; Strings—Salt Licenses, Preservation Contracts, Trade Routes.
\end{itemize}
Special Rule — Iron Haven Privilege: Ships flying no nation's colors are treated as neutral parties. Smuggle starts Controlled; Expose a pirate's identity starts Desperate. \\
Adventure Starters: A rival fleet's flagship appears in harbor with no crew • The Iron Tithe ledgers show impossible debts • A ghost ship is sinking legitimate vessels • The Harbor Master's Court declares martial law.

\sectionrule

\section*{8) Player-Facing Options (light)}
Urban Talents: Streetwise, Ledger-Eye, Bridge-Runner, Rite-Speaker, Patrol-Friend, Crowd-Reader, Shipwright, Corsair, Harbor Pilot, Salt-Trader. \\
Background Hooks: Guild-taught, Pilot's Kin, Temple Page, Festival Barker, Caravan Clerk, Exarchate Apprentice, Ship's Boy, Corsair Captain, Harbor Guard, Salt Miner, River Runner, Dock Master. \\
Temptations (optional): Replace vice with Ambition / Comfort / Fury / Pride. Track a 4-segment Temptation Clock that the GM can press in districts tagged with that flavor; spend Boons to resist or clear.

\sectionrule

\section*{9) GM Toolkit}
Prep: City, dials, factions (4), volatile stakes (2), districts (6--8), first draw (Spade/Heart/Club), Situation/Counter clocks. \\
SB Spends: faux pas • leverage leak • crowd turns • permit pulled • patrol biased • festival crush • toll chain drops • ship quarantine • cargo inspection • guild audit • noble scandal • temple censure • watch curfew • harbor closure • price fixing • counterfeit goods • debt collector • bounty hunter • informer betrayal • rival sabotage. \\
Default Ripple: Success → ally Influence +1 (or +2 if on Aim); enemy Exposure +1; clean win reduces ally Exposure --1. Failure → enemy Influence +1 (or +2 on Aim); ally Exposure +1; internal failure also Stability +1. \\
DV Cheats: Petition 2--4 (temple/court), Broker 3--4 (neutral ground helps), Expose 3--5 (proof lowers), Infiltrate 3--5 (wards raise). \\
\textbf{Three-Clock Rule:} Only maintain three active clocks at once—Situation, Counter-Clock, and Campaign. Merge or retire others.

\subsection*{5-Minute City Prep:}
\begin{enumerate}
    \item Pick 1 City Kit
    \item Circle 4 Factions
    \item Highlight 2 Districts for scene
    \item Note 1 Special Rule
    \item Draw 1 Urban Deck card
    \item = Ready for first scene
\end{enumerate}

\sectionrule

\section*{10) Printable Aids}
City Sheet • District Cards (tags/effects) • Factions \& Strings minis • Urban Deck prompts • Chase Ladders sidecard • Events \& Omens mini-table • Investigation Tracker • Relationship Map Template.

\sectionrule

\section*{Changelog}
v0.4 — Final refinement pass; enhanced Zakov city kit with detailed lore; expanded district tags; clarified string cutting examples; added investigation clock guidance; expanded social SB menu; reinforced three-clock rule; added urban action synergy notes; enhanced chase complications; expanded events/omens; added prep tools for GM.

\textbf{End of v0.4}

\end{document}