hor{}
\date{}

\begin{document}
\maketitle

\section{Preamble}
There are places where the world forgets its laws.
Doorways that no one admits to knocking upon.
Names that rewrite the one who dares to speak them.
Shadows are not evil−−−they are consequences given shape. They are the costs we refuse to tally, the truths we bury, the debts that return to collect themselves.

\begin{RuleBox}
\Term{Witchcraft} is reciprocal and intimate. A working binds the witch to something that can answer back: a threshold, a spirit, a vow, a Name, a patron, a memory, a debt.
\end{RuleBox}

\section{Core Principles}
\subsection{Shadows}
Shadows arise wherever truth is denied, debt is unacknowledged, or identity is fractured by unmet need. They are narrative pressure given teet\documentclass[11pt,letterpaper]{article}
\usepackage[margin=0.85in]{geometry}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{xcolor}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage[dvipsnames]{xcolor}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\usepackage{fancyhdr}

% Formatting
\pagestyle{fancy}
\fancyhf{}
\renewcommand{ℎeadrulewidth}{0pt}
\rfoot{þepage}

% Section styling
\titleformat{\section}
  {\normalfontŁarge\bfseries\scshape}{þesection}{1em}{}
\titleformat{\subsection}
  {\normalfontłarge\bfseries}{þesubsection}{1em}{}
\titleformat{⫕section}
  {\normalfont\normalsize\bfseriesℹtshape}{þesubsubsection}{1em}{}

% Custom boxes for mechanics
≠wtcolorbox{ritestat}{
  colback=gray!5,
  colframe=gray!50!black,
  boxrule=0.5pt,
  arc=2mm,
  left=6pt,
  right=6pt,
  top=4pt,
  bottom=4pt,
  fonttitle=\bfseries,
  title={Rite Parameters}
}

≠wtcolorbox{mechanic}[1][]{
  colback=white,
  colframe=black!70,
  boxrule=0.4pt,
  arc=1mm,
  left=5pt,
  right=5pt,
  fonttitle=\bfseries\scshape\small,
  title={#1}
}

≠wtcolorbox{sidebar}{
  colback=black!3,
  colframe=black!40,
  boxrule=0.4pt,
  arc=1mm,
  left=6pt,
  right=6pt,
  top=4pt,
  bottom=4pt
}

% Game terms
≠wcommand{øbligation}[1]{\textsc{Obligation~#1}}
≠wcommand{\dv}[1]{\textsc{DV}~#1}
≠wcommand{\xp}[1]{#1~\textsc{xp}}
≠wcommand{\sbmark}{\textsc{sb}}


\usepackage{hyperref}
ℎypersetup{colorlinks=true, linkcolor=black, urlcolor=blue}

\setlist[itemize]{leftmargin=1.25em}
\setlist[enumerate]{leftmargin=1.6em}

≠wcommand{\Term}[1]{\textbf{#1}}
≠wcommand{\DV}{\Term{DV}}
≠wcommand{§B}{\Term{SB}}

≠wtcolorbox{RuleBox}[1][]{colback=black!2,colframe=black!60,arc=2mm,boxrule=0.6pt,left=1.2mm,right=1.2mm,top=1.0mm,bottom=1.0mm,#1}
≠wtcolorbox{RiteBox}[1][]{colback=black!1,colframe=black!35,arc=2mm,boxrule=0.5pt,left=1.2mm,right=1.2mm,top=1.0mm,bottom=1.0mm,#1}

\title{The Book of Shadows}
\auth.

\subsection{Thresholds}
A threshold is any place where a state changes: waking to dreaming, innocence to agency, belonging to exile, life to echo. Crossing invites witness−−−and witnesses demand accounting.

\subsection{Name−Truth}
A \Term{Name} is not a label. It is a claim with weight. Names can be protected, traded, buried, sharpened, dulled, or stolen. When a working touches a Name, it touches the self behind it.

\section{Witchcraft Procedure}
\subsection{Casting a Working}
When you attempt a working, state the \Term{Intent}, the \Term{Vector}, and the \Term{Price}.

\begin{itemize}
  ℹtem \Term{Intent}: what changes (veil, bind, unmake, reveal, curse, sever, restore).
  ℹtem \Term{Vector}: what carries the change (name, oath, threshold, shadow, dream, debt, token, bloodless offering).
  ℹtem \Term{Price}: what you pay (Shadow, Shame, Identity).
\end{itemize}

Then roll \Term{Attribute + Witchcraft} vs \DV{} set by scope.

\subsection{Scope and Difficulty}
\begin{center}
\begin{tabularx}{0.92łinewidth}{@{}lXc@{}}
⊤rule
\Term{Scope} & \Term{Examples} & \DV \\
∣rule
Subtle / Personal & blur attention, soften a memory−edge, silence a doorway, hide a token & 2 \\
Situational & bind a promise, distort a reflection, mark a Name, hinder pursuit & 3 \\
Scene−Shaping & compel a confession, veil a crowd, turn a gathering into a threshold & 4 \\
Identity−Altering & rewrite a role, bury a name−right, sever a bond, harvest an echo & 5+ \\
⊥tomrule
\end{tabularx}
\end{center}

\subsection{Outcomes}
\begin{itemize}
  ℹtem \Term{Full Success}: The working takes hold as intended. Pay the chosen price.
  ℹtem \Term{Partial Success}: The working takes hold with a complication. Pay the price and the GM adds a consequence.
  ℹtem \Term{Miss}: The working fails or twists. Pay the price and the GM escalates the consequence.
\end{itemize}

\subsection{Push It}
Before the roll or after a partial success, you may \Term{Push It} to increase scope or intensity.
Choose one:
\begin{itemize}
  ℹtem Mark \Term{+1 Shadow} and gain \Term{+1d}.
  ℹtem Mark \Term{+1 Shame} and gain \Term{+1 Effect}.
  ℹtem Mark \Term{+1 Identity Strain} and reduce the \DV{} by 1 (minimum 2).
\end{itemize}

\section{Prices}
Every working demands a price. Choose \Term{one} unless a rite specifies otherwise.

\subsection{Shadow}
\begin{itemize}
  ℹtem You accrue spiritual residue and narrative pressure.
  ℹtem The GM may introduce shadow−tells: cold glass, wrong reflections, trailing whispers, predatory coincidences.
\end{itemize}

\subsection{Shame}
\begin{itemize}
  ℹtem You accept an internal consequence: a Condition, a vow−compulsion, a visible hesitation, a ritual aversion.
  ℹtem The GM may invoke Shame when your past choice becomes socially or spiritually relevant.
\end{itemize}

\subsection{Identity}
\begin{itemize}
  ℹtem You relinquish a piece of self: a cherished memory dulled, a role surrendered, a name−right compromised, a bond weakened.
  ℹtem Identity prices always change what the character can say with full certainty about who they are.
\end{itemize}

\section{Identity Strain}
Identity Strain tracks cumulative erosion from rites that touch the self directly.

\begin{RuleBox}
\Term{Identity Strain} is a 4−step track. Mark it when a working severs, rewrites, harvests, or binds a Name; when you Push It with Identity; or when a rite demands it.
\end{RuleBox}

\begin{itemize}
  ℹtem At \Term{4}, you must resolve strain at the end of the scene: roll \Term{Spirit + Resolve} vs \DV{} 4.
  ℹtem \Term{Success}: Clear 1 Identity Strain.
  ℹtem \Term{Failure}: The GM chooses or offers a hard bargain: a memory blurs, a bond becomes conditional, a Name becomes \emph{buried}, or an echo−self becomes active.
\end{itemize}

\section{Observances}
Observances are the lived rules witches follow to keep magic from turning inward and hollowing them out.

\subsection{Common Observances}
\begin{itemize}
  ℹtem \Term{Count your debts aloud} once per week at a threshold you respect.
  ℹtem \Term{Do not cross the same threshold twice in one night} unless invited.
  ℹtem \Term{Never speak a Name in anger} if you intend to keep it.
  ℹtem \Term{Leave an offering} when a working succeeds cleanly: a coin, a braid of grass, a thread, a pinch of ash.
  ℹtem \Term{Do not take from the dead without a witness}. The witness may be a spirit, a person, or the threshold itself.
\end{itemize}

\subsection{Order Observances}
⫕section{The Shed Vigil (Lethai−ar of Isoka)}
\begin{RiteBox}
\Term{Signs}\\
Rings of shed bone or horn; cuts on wrists marked with ash; small piles of discarded tokens at crossroads; masks split vertically, bound by cord.
\medskip

\Term{Etiquette}\\
\begin{itemize}
  ℹtem Do not refuse a witch's request for a symbolic offering.
  ℹtem Do not speak someone's \Term{buried name} aloud.
  ℹtem Accept the loss you cannot hide−−−loss defines you.
  ℹtem Never interrupt a Severance Rite.
\end{itemize}
\end{RiteBox}

⫕section{Covenant Observances (Lunera)}
\begin{RiteBox}
\Term{Bargains and Prices}\\
\begin{itemize}
  ℹtem \Term{Memory Tithe}: Offer a cherished memory−−−not erased, but dulled for you and sharpened for another.
  ℹtem \Term{Hidden Truth}: Reveal a secret that would transform someone's understanding of their past.
  ℹtem \Term{Kept Silence}: Maintain a night of absolute silence at a chosen threshold; break it and the next omen arrives twisted.
  ℹtem \Term{Selective Sight}: In exchange for perfect clarity in one specific matter, accept permanent blind spots elsewhere (agree on 1−−2 narrative blind spots).
\end{itemize}
\end{RiteBox}

\section{Universal Workings}
These are patterns of craft shared across Orders; the difference is in doctrine, taboos, and cost.

\subsection{Veil Working}
Hide, blur, silence, soften, misdirect.
\begin{itemize}
  ℹtem On success, gain \Term{+1d} on the next stealth, deception, or evasion roll in the scene.
  ℹtem On partial, the veil works but leaves a tell: a reflection moves late, a shadow points wrong, a witness feels watched.
\end{itemize}

\subsection{Binding Working}
Hold, anchor, compel, constrain, mark.
\begin{itemize}
  ℹtem On success, impose a constraint that must be narrated and testable (``cannot speak the lie,'' ``cannot cross the doorway,'' ``must answer the question,'' etc.).
  ℹtem On partial, the constraint holds but spawns a counter−pressure (a timer, a spirit notice, a rival claim).
\end{itemize}

\subsection{Unmaking Working}
Break oaths, dissolve lesser magic, strip labels, undo a threshold knot.
\begin{itemize}
  ℹtem On success, remove one active magical effect, one binding tag, or one scene−level occult constraint.
  ℹtem On miss, something else unravels instead (GM choice): an unrelated oath frays, a Name loosens, a memory bleeds into the present.
\end{itemize}

\subsection{Crisis Working}
Forbidden intensity. The working changes the entire scene.
\begin{itemize}
  ℹtem Mark \Term{+2 Identity Strain}.
  ℹtem On success, the scene changes as declared.
  ℹtem On partial, it changes−−−and a second consequence arrives with it.
  ℹtem On miss, the working becomes a hungry thing: it demands resolution through a timer, a bargain, or a witness.
\end{itemize}

\section{Selected Rites}
Rites are learned patterns with fixed doctrine. Each rite lists its tag, scope, and requirements.

\subsection{Rite of the Ashen Knife ℎfill \textnormal{[SEVER] Low, 4 XP}}
\begin{RiteBox}
\Term{Type}: Action  
\Term{Materials}: Ash; a token willingly surrendered \\
\Term{Effect}: Create a knife of ash and intent for the scene.
\begin{itemize}
  ℹtem Gain \Term{+1 Position} when breaking bindings.
  ℹtem On a success, you may symbolically ``cut'' a minor hindrance or complication.
\end{itemize}
\Term{Price}: Shadow or Identity. \\
\Term{Push It}: Mark \Term{+1 Identity Strain} to cut a more meaningful constraint (GM sets boundary).
\end{RiteBox}

\subsection{Rite of Echoed Dreams ℎfill \textnormal{[ECHO] Standard, 7 XP}}
\begin{RiteBox}
\Term{Type}: Action; Close  
\Term{Materials}: Shared token from a past moment (locket, scrap of letter, shard of glass) \\
\Term{Effect}: Draw out a vivid waking−dream from a willing subject. Ask up to two:
\begin{itemize}
  ℹtem ``Who do you most fear becoming?''
  ℹtem ``What moment do you wish you could change?''
  ℹtem ``What truth do you not dare to look at?''
\end{itemize}
The dream answers in symbols; the GM ties each symbol to a concrete NPC, place, or timer. You gain \Term{+1d} when acting with compassion or leverage on that truth this session. \\
\Term{Push It}: You may enter the dream with them; both gain \Term{+1d} to resist fear or confusion linked to that memory this session, and you mark \Term{+1 Obligation} to the relevant patron or witness.
\end{RiteBox}

\subsection{Hex of the Clinging Shadow ℎfill \textnormal{[HEX] Standard, 8 XP}}
\begin{RiteBox}
\Term{Type}: Scene; Near  
\Term{Materials}: Fragment of the target's reflection (hair from their comb, a sketch of their face, etc.) \\
\Term{Effect}: Call a second shadow to cling to the target for the scene (or until dispelled).
\begin{itemize}
  ℹtem They cannot fully hide their intentions; social deception rolls suffer \Term{−1d}.
  ℹtem When they attempt betrayal or flee, their shadow lags or points the wrong way (GM adds a tell or complication).
\end{itemize}
\Term{Push It}: Bind the shadow to a \Term{Shadow's Debt [4]} timer. When it fills (through lies, harm, cowardice), the shadow manifests as an echo−spirit that can testify or act once on your whispered command.
\end{RiteBox}

\subsection{Curse of the Unanswered Reflection ℎfill \textnormal{[CURSE] High, 10 XP}}
\begin{RiteBox}
\Term{Type}: Long; Far  
\Term{Materials}: A mirror cracked in three pieces, each named for a truth the target refuses \\
\Term{Effect}: Mark a \Term{Refusal [6]} timer on the target (individual or faction). Each time they double down on denial, cruelty, or willful ignorance, advance the timer.
\begin{itemize}
  ℹtem At 3/6: distorted reflections whisper failings (GM adds fear, doubt, paranoia as Conditions).
  ℹtem At 6/6: a Reckoning Scene is forced: confrontation, trial, or nightmare where the hidden truth is dragged into the open.
\end{itemize}
\Term{Price}: Identity (mandatory). Mark \Term{+1 Identity Strain}.
\end{RiteBox}

\section{Suppression and Witness}
Witchcraft hates clarity. When a working is cast under scrutiny, doctrine, lanternlight, or formal correction, increase \DV{} by 1. On a miss (or on a complication), the GM may generate §B{} toward reprisal, exposure, or a hunter response.

\section{Quick Reference}
\begin{RuleBox}
\begin{itemize}
  ℹtem Declare \Term{Intent}, \Term{Vector}, \Term{Price}.
  ℹtem Roll \Term{Attribute + Witchcraft} vs scope \DV{} (2−−5+).
  ℹtem Pay \Term{one Price}: Shadow, Shame, or Identity.
  ℹtem Optional: \Term{Push It} for +1d / +1 Effect / \DV{} −1 (at a cost).
  ℹtem Track \Term{Identity Strain [4]} for self−touching magic.
\end{itemize}
\end{RuleBox}

\section{Witch Orders and Their Rites}

Witchcraft is not unified. Each Order encodes doctrine, taboo, and price into its rites, shaping how power is expressed and what it consumes in return. To join an Order is to accept its view of what must be sacrificed for magic to work.

\subsection{The Night Court of Mab}
The Night Court teaches that beauty is leverage, promises are weapons, and truth is negotiable under moonlight. Mab's witches defend ambiguity and punish certainty.

⫕section*{Doctrine}
Reality bends most easily where desire and fear overlap. An oath half−kept is more powerful than one fulfilled.

⫕section*{Observances}
\begin{itemize}
  ℹtem Never speak an unvarnished truth beneath the open moon.
  ℹtem All bargains must contain a flaw, spoken or hidden.
  ℹtem Broken promises are never wasted.
\end{itemize}

⫕section*{Signature Rites}

\paragraph{Thorn of Oaths−Broken ℎfill \textnormal{[HEX] Standard}}
Impose a spiritual wound on a target who has broken a vow.
\begin{itemize}
  ℹtem The target suffers \emph{Oath−Bleed}: Position is reduced by 1 when acting against their word.
  ℹtem On a miss, the witch inherits the recoil of the broken oath.
\end{itemize}
\Term{Price}: Shame or Shadow.

\paragraph{Moon−Veil Compact ℎfill \textnormal{[VEIL] Low}}
Wrap a subject in fae ambiguity.
\begin{itemize}
  ℹtem Gain +1d to deception or concealment for the scene.
  ℹtem Leaves a glamour−tell visible only in reflections.
\end{itemize}
\Term{Price}: Shadow.

\subsection{The Ash Hags of Morag}
Morag's witches embody rot, endings, and consequence. They do not forgive; they conclude.

⫕section*{Doctrine}
Everything ends. Magic exists to make sure it ends \emph{properly}.

⫕section*{Observances}
\begin{itemize}
  ℹtem Never reverse a punishment once begun.
  ℹtem Bones are records; treat them with respect.
  ℹtem Mercy is a lie told by those who fear endings.
\end{itemize}

⫕section*{Signature Rites}

\paragraph{Bone−Truth Casting ℎfill \textnormal{[REVEAL] Standard}}
Read consequence in bone or ash.
\begin{itemize}
  ℹtem Ask one question about an unavoidable outcome.
  ℹtem The answer cannot be softened.
\end{itemize}
\Term{Price}: Shame.

\paragraph{Curse of Settled Ash ℎfill \textnormal{[CURSE] High}}
Mark a target for inevitable decline.
\begin{itemize}
  ℹtem Create a \emph{Decay [6]} timer that advances when the target resists accountability.
  ℹtem At completion, a final reckoning scene occurs.
\end{itemize}
\Term{Price}: Identity (mandatory).

\subsection{The Handmaidens of Livaea}
Livaea's witches rule through attraction, influence, and emotional architecture. Power flows through wanting.

⫕section*{Doctrine}
No one is free of desire. Control what is desired, and you control the world.

⫕section*{Observances}
\begin{itemize}
  ℹtem Never refuse a sincere confession.
  ℹtem Desire must always be mirrored, never denied.
  ℹtem Public humiliation is forbidden; private leverage is sacred.
\end{itemize}

⫕section*{Signature Rites}

\paragraph{Thread of Longing ℎfill \textnormal{[BIND] Standard}}
Bind two subjects through unfulfilled desire.
\begin{itemize}
  ℹtem Each gains leverage over the other until the desire is resolved or destroyed.
\end{itemize}
\Term{Price}: Shame.

\paragraph{Velvet Mirror ℎfill \textnormal{[REVEAL] Low}}
Reveal what a subject wishes they did not want.
\begin{itemize}
  ℹtem Gain +1 Position when acting on this knowledge this session.
\end{itemize}
\Term{Price}: Shadow.

\subsection{The Rainmaidens of Raéyn}
Raéyn's witches are tempestuous elementalists whose emotions shape storm and tide.

⫕section*{Doctrine}
Emotion is not weakness. It is weather.

⫕section*{Observances}
\begin{itemize}
  ℹtem Never deny your mood.
  ℹtem Water must be returned to water.
  ℹtem Storms demand witnesses.
\end{itemize}

⫕section*{Signature Rites}

\paragraph{Tidebreaker Surge ℎfill \textnormal{[STORM] Standard}}
Redirect momentum in a conflict.
\begin{itemize}
  ℹtem Allies gain +1 Position; enemies suffer −1d for one exchange.
\end{itemize}
\Term{Price}: Shadow.

\paragraph{Mist−Crowned Omen ℎfill \textnormal{[REVEAL] Low}}
Read fate in condensation and rainfall.
\begin{itemize}
  ℹtem Ask one question about imminent danger or opportunity.
\end{itemize}
\Term{Price}: Shame.

\subsection{The Thorns of Malachai}
The Thorns are not servants of a patron. They are adherents to a principle: identity is harvestable. What is discarded can be claimed, reshaped, and weaponized.

⫕section*{Doctrine}
What the world refuses to hold must be taken. Forgotten selves are resources.

⫕section*{Observances}
\begin{itemize}
  ℹtem Never destroy an identity fragment while it still remembers pain.
  ℹtem Names are extracted, not spoken.
  ℹtem Witnesses are liabilities.
\end{itemize}

⫕section*{Signature Rites}

\paragraph{Shadow Extraction ℎfill \textnormal{[SEVER] Standard}}
Strip a fragment of identity from a target.
\begin{itemize}
  ℹtem Mark \emph{Identity Strain} on the target.
  ℹtem Remove or distort a memory, bond, or role for the scene or longer.
\end{itemize}
\Term{Price}: Identity (mandatory).

\paragraph{False Echo Weaving ℎfill \textnormal{[ECHO] High}}
Create a shadow−double seeded with stolen memory.
\begin{itemize}
  ℹtem The echo may testify, mislead, or act once before unraveling.
\end{itemize}
\Term{Price}: Identity and Shadow.

\paragraph{Harvest of the Unclaimed ℎfill \textnormal{[RITUAL] High}}
Ritualize mass forgetting.
\begin{itemize}
  ℹtem Advance a shared \emph{Identity Collapse} timer in the area.
  ℹtem At completion, the Thorns gain a permanent asset or manifestation.
\end{itemize}
\Term{Price}: Identity Strain for all participants.

\section{On Familiars}
A familiar is not a pet, a tool, or a summoned thing.
A familiar is a \emph{shared consequence}.

It is formed when a witch externalizes part of their shadow, memory, instinct, or unspoken need into a living conduit that can act, witness, and answer back. Once bound, neither witch nor familiar remains whole in the way they were before.

\begin{RuleBox}
A \Term{Familiar} is a persistent bond that allows a witch to relocate risk, perception, and cost. It amplifies magic by distributing consequence−−−but never removes it.
\end{RuleBox}

\section{Creating a Familiar}
To create a familiar, a witch performs a binding at a meaningful threshold and commits to an ongoing obligation.

\subsection{Requirements}
\begin{itemize}
  ℹtem A willing or capturable vessel (animal, spirit, echo, object−entity)
  ℹtem A threshold significant to the witch
  ℹtem A personal offering that cannot be reclaimed
\end{itemize}

\subsection{The Binding}
State:
\begin{itemize}
  ℹtem \Term{What the Familiar Knows} (instinct, secret, fear, hunger)
  ℹtem \Term{What the Familiar Wants} (safety, sensation, purpose, attention)
  ℹtem \Term{What the Witch Gave Up} (memory, role, bond, certainty)
\end{itemize}

Roll \Term{Spirit + Witchcraft} vs \DV{} 4.

\begin{itemize}
  ℹtem \Term{Success}: The bond forms cleanly.
  ℹtem \Term{Partial}: The bond forms with a flaw (GM defines a behavioral complication).
  ℹtem \Term{Miss}: The bond forms−−−but the familiar defines the terms of attention, need, or return.
\end{itemize}

Mark \Term{+1 Identity Strain}. This mark cannot be avoided.

\section{Familiar Nature}
Choose or define the familiar's nature. This shapes how it interacts with magic and consequence.

\begin{center}
\begin{tabularx}{0.95łinewidth}{@{}lX@{}}
⊤rule
\Term{Nature} & \Term{Expression} \\
∣rule
Animal & Heightened senses, instinctual reactions, physical presence \\
Spirit & Insubstantial movement, omen−bearing, threshold traversal \\
Echo & Memory replay, voice mimicry, emotional residue \\
Construct & Inscription, containment, durability, symbolic action \\
Parasite & Hunger−driven aid, escalation under stress \\
⊥tomrule
\end{tabularx}
\end{center}

\section{Bond Track}
Each familiar has a \Term{Bond Track [4]} shared with its witch.

Mark Bond when:
\begin{itemize}
  ℹtem The familiar assists in a working
  ℹtem The familiar witnesses violence, betrayal, or confession
  ℹtem The witch Pushes a working through the familiar
\end{itemize}

When the Bond Track fills, choose one:
\begin{itemize}
  ℹtem The familiar demands fulfillment (food, secrecy, rest, release)
  ℹtem The familiar acts independently in line with its nature
  ℹtem The bond destabilizes; the GM introduces a rupture or misalignment
\end{itemize}

Clear Bond only by honoring the familiar's needs.

\section{Using a Familiar}
A familiar may be invoked once per scene without a roll for minor assistance consistent with its nature.

For greater effect, choose one option and roll \Term{Attribute + Witchcraft}:

\subsection{Extend Perception}
See, hear, smell, or sense through the familiar.
\begin{itemize}
  ℹtem Gain \Term{+1d} on awareness or tracking
  ℹtem Cost: Mark Bond
\end{itemize}

\subsection{Relay a Working}
Cast a working as if you were present where the familiar is.
\begin{itemize}
  ℹtem Increase \DV{} by 1
  ℹtem Cost: Familiar shares all backlash
\end{itemize}

\subsection{Absorb Consequence}
Redirect fallout from a working.
\begin{itemize}
  ℹtem Convert 1 Shadow or Shame into 1 Bond
  ℹtem On a miss, the familiar manifests the consequence physically or socially
\end{itemize}

\section{Familiars and Cost}
Familiars do not erase price−−−they \emph{translate} it.

\subsection{Shadow Through the Familiar}
Shadow may appear as:
\begin{itemize}
  ℹtem Unnatural behavior
  ℹtem Aggression or withdrawal
  ℹtem Physical mutation or spectral distortion
\end{itemize}

\subsection{Shame Through the Familiar}
Shame may cause:
\begin{itemize}
  ℹtem Avoidance of certain people or places
  ℹtem Vocalizations or gestures at inopportune times
  ℹtem Involuntary displays of guilt or fear
\end{itemize}

\subsection{Identity Through the Familiar}
Identity loss may:
\begin{itemize}
  ℹtem Bind the familiar more tightly than intended
  ℹtem Transfer a trait, memory, or name−fragment
  ℹtem Cause the familiar to speak truths the witch cannot
\end{itemize}

\section{Harm to Familiars}
Familiars cannot be safely replaced.

When a familiar is harmed:
\begin{itemize}
  ℹtem The witch immediately marks \Term{+1 Identity Strain}
  ℹtem The witch rolls \Term{Spirit + Resolve} vs \DV{} 3
\end{itemize}

On failure, gain a lingering Condition tied to the harm.

If a familiar is destroyed:
\begin{itemize}
  ℹtem The witch permanently loses something bound during creation
  ℹtem All Bond clears
  ℹtem The witch gains a recurring omen, scar, or echo
\end{itemize}

\section{Observances of Familiar−Binding}
\subsection{Common Observances}
\begin{itemize}
  ℹtem Never force a familiar to cross a threshold you would refuse yourself.
  ℹtem Do not lie to your familiar; silence is acceptable, deception is not.
  ℹtem Feed the familiar before feeding yourself after a working.
  ℹtem Do not bind more than one familiar without release or death between.
\end{itemize}

\subsection{Signs of a Failing Bond}
\begin{itemize}
  ℹtem The familiar avoids eye contact or reflection
  ℹtem Shared dreams fracture or repeat
  ℹtem The familiar reacts before the witch notices danger
  ℹtem Names spoken near the familiar echo incorrectly
\end{itemize}

\section{Selected Familiar Rites}

\subsection{Rite of the Watching Shadow ℎfill \textnormal{[BIND] Low, 4 XP}}
\begin{RiteBox}
\Term{Effect}: Bind a minor shadow or animal as a familiar of vigilance.
\begin{itemize}
  ℹtem Once per session, the familiar may warn of imminent danger without a roll.
\end{itemize}
\Term{Cost}: Shadow.
\end{RiteBox}

\subsection{Rite of Shared Breath ℎfill \textnormal{[LINK] Standard, 7 XP}}
\begin{RiteBox}
\Term{Effect}: Share physical strain between witch and familiar for the scene.
\begin{itemize}
  ℹtem Split Fatigue or Harm as chosen.
\end{itemize}
\Term{Cost}: Shame or Identity.
\end{RiteBox}

\subsection{Rite of the Last Witness ℎfill \textnormal{[SEVER] High, 10 XP}}
\begin{RiteBox}
\Term{Effect}: If the witch would fall or lose consciousness, the familiar may act once to preserve a truth, message, or memory.
\begin{itemize}
  ℹtem Afterward, the bond immediately destabilizes.
\end{itemize}
\Term{Cost}: Identity (mandatory).
\end{RiteBox}

\section{Quick Reference}
\begin{RuleBox}
\begin{itemize}
  ℹtem Familiars relocate consequence; they do not remove it.
  ℹtem Creating a familiar always marks \Term{Identity Strain}.
  ℹtem Track \Term{Bond [4]} for escalating tension.
  ℹtem Harm to a familiar harms the witch.
  ℹtem A destroyed familiar leaves a permanent absence.
\end{itemize}
\end{RuleBox}

\section{Witchcraft Followers}

In the world of Witchcraft, followers are not merely assistants−−−they are the living conduits of power, the silent witnesses to hidden truths, and the keepers of sacred obligations. A Witchcraft follower exists at the threshold between the mundane and the magical, bound by pacts, oaths, and the ever−present weight of consequence.

\subsection{The Nature of Witchcraft Followers}

Witchcraft followers are those who have willingly or unwillingly bound themselves to the forces of the Witchcraft system. Unlike ordinary allies, these followers exist in the liminal space where the world's hidden currents flow most strongly. They are not merely helpers but extensions of the witch's pact, bearing the weight of Obligation, the touch of Shadow, and the ever−present threat of the Unanswered Reflection.

\subsection{Follower Types in Witchcraft}

⫕section{The Oathsworn}
Born from the moment a follower swears a sacred promise to a Patron, the Oathsworn are the most common form of Witchcraft follower.

\textbf{Characteristics:}
\begin{itemize}
    ℹtem Carries a visible mark of their oath (a sigil, scar, or token)
    ℹtem Can sense when the oath is in peril
    ℹtem May speak the true name of the Patron when called upon
\end{itemize}

\textbf{Magical Condition States:}
\begin{itemize}
    ℹtem \textbf{True−Bound}: Oath is strong; +1 die to actions related to the Patron's domain
    ℹtem \textbf{Fractured}: Oath is strained; −1 die to Patron−related actions; +1 Obligation segment
    ℹtem \textbf{Shattered}: Oath is broken; follower becomes a Wraith, seeking to make the witch pay
\end{itemize}

⫕section{The Threshold−Keepers}
These followers walk the edge between worlds, maintaining the boundaries that keep the witch's power contained.

\textbf{Characteristics:}
\begin{itemize}
    ℹtem Can sense and reinforce magical boundaries
    ℹtem May temporarily lower a Ward for the witch
    ℹtem Drawn to crossroads, doorways, and places of transition
\end{itemize}

\textbf{Magical Condition States:}
\begin{itemize}
    ℹtem \textbf{Ward−Safe}: Threshold connection strong; +1 die to boundary−related actions
    ℹtem \textbf{Frayed}: Threshold connection weakening; −1 die to boundary actions; +1 Exposure on threshold use
    ℹtem \textbf{Breached}: Threshold connection broken; may attract unwanted attention from beyond
\end{itemize}

⫕section{The Shadow−Speaking}
Bound to the very essence of Shadow, these followers can perceive the unspoken truths that haunt the witch's path.

\textbf{Characteristics:}
\begin{itemize}
    ℹtem Sees the shadows that follow the witch
    ℹtem Can speak the words that should remain unspoken
    ℹtem Haunted by the things they've witnessed
\end{itemize}

\textbf{Magical Condition States:}
\begin{itemize}
    ℹtem \textbf{Clear−Sighted}: Shadow−sight is strong; +1 die to insight/truth−detection
    ℹtem \textbf{Clouded}: Shadow−sight is weakening; −1 die to insight/truth−detection; +1 Exposure on truth−seeing
    ℹtem \textbf{Shrouded}: Shadow−sight is lost; may see false shadows or be consumed by darkness
\end{itemize}

\subsection{Universal Witchcraft Follower Mechanics}

⫕section{Obligation Tracking}
Witchcraft followers have their own Obligation timer that is tied to the witch's pact.

\begin{itemize}
    ℹtem \textbf{Obligation Timer }: 6 segments (tied to the witch's Obligation Capacity)
    ℹtem \textbf{Filling the Timer }: When the follower's timer fills:
    \begin{itemize}
        ℹtem 3 segments: The follower gains +1 die to actions related to the pact
        ℹtem 6 segments: A demand from the Patron is made of the witch through the follower
    \end{itemize}
    ℹtem \textbf{Clearing Obligation}: The witch must perform a service for the Patron to reduce the follower's Obligation timer
\end{itemize}

⫕section{Shadow Exposure}
Witchcraft followers accumulate Shadow Exposure from their proximity to the witch's power.

\begin{itemize}
    ℹtem \textbf{Exposure Effects:}
    \begin{itemize}
        ℹtem 1: Can see the witch's Shadow as a faint aura
        ℹtem 2: Hears the whispers of the Unanswered Reflection
        ℹtem 3: Overexposed − Suffers Fatigue 1, −1 die to all rolls, and may trigger a minor Shadow event
    \end{itemize}
\end{itemize}

⫕section{Shadow Harm}
Magical followers track Shadow Harm, which represents damage to their connection with the shadow world.

\begin{itemize}
    ℹtem \textbf{Harm Effects:}
    \begin{itemize}
        ℹtem 1: Shadow world begins to bleed into their reality
        ℹtem 2: Suffers −1 die to social rolls with those who cannot see the Shadow
        ℹtem 3: Dissonant − Cannot use their magical abilities; may attract Shadow entities
    \end{itemize}
\end{itemize}

⫕section{Magical Assistance}
Witchcraft followers can assist in witchcraft−related actions:

\begin{itemize}
    ℹtem \textbf{Standard Assistance (1 Boon cost):}
    \begin{itemize}
        ℹtem Grant +1 die to witchcraft rolls within their specialty
        ℹtem Share their insight about the ritual or Patrons
    \end{itemize}
    ℹtem \textbf{Ritual Assistance (2 Boon cost):}
    \begin{itemize}
        ℹtem Let the witch borrow their connection for one roll
        ℹtem Stabilize a ritual that is about to go awry
        ℹtem Reveal a hidden truth about the ritual's requirements
    \end{itemize}
\end{itemize}

\subsection{How to Gain a Witchcraft Follower}

Witchcraft followers are not easily acquired. They require:

\begin{itemize}
    ℹtem A significant story moment that creates the pact or bond
    ℹtem A clear oath or commitment between the witch and the follower
    ℹtem A narrative reason for the follower to have their connection to the Witchcraft system
\end{itemize}

Gaining a Witchcraft follower costs 4 XP, but comes with an immediate 2 Obligation segments that the witch must bear.

\subsection{The Cost of Connection}

Witchcraft followers are not mere tools−−−they carry the weight of the witch's choices. When a follower falls, it's not just a loss of an ally−−−it's a fracture in the pact itself.

\begin{itemize}
    ℹtem When a follower dies while in service, the witch gains 3 Obligation segments
    ℹtem The witch's next ritual has a −1 die penalty (the weight of loss)
    ℹtem The follower's spirit may return as a Shadow entity, seeking resolution
\end{itemize}

\subsection{Example: The Oathsworn Follower}

\textbf{Name:} Anya the Steadfast \\
\textbf{Patron:} The Gallow's Bell \\
\textbf{Oath Mark:} A small bell−shaped scar on her left palm \\
\textbf{Starting Condition:} True−Bound \\

\textbf{Oathsworn Condition State Table:}
\begin{itemize}
    ℹtem \textbf{True−Bound}: +1 die to actions related to the Gallow's Bell's domain (truth, justice)
    ℹtem \textbf{Fractured}: −1 die to Patron−related actions; +1 Obligation segment
    ℹtem \textbf{Shattered}: Follower becomes a Wraith, seeking to make the witch answer for their broken oath
\end{itemize}

\textbf{Oathsworn Special Actions:}
\begin{itemize}
    ℹtem \textbf{Oath−Call:} Once per scene, Anya can call on the Gallow's Bell to verify a statement; the speaker must speak truth or suffer 1 Harm
    ℹtem \textbf{Ward−Safe:} Once per ritual, she can stabilize the ritual space, reducing the risk of Backlash by one level
\end{itemize}

\textbf{Oathsworn Shadow Exposure Effects:}
\begin{itemize}
    ℹtem 1: Sees the faint glow of truth in the air
    ℹtem 2: Hears the echoes of unspoken promises
    ℹtem 3: Overexposed − Suffers Fatigue 1, −1 die to all rolls, and may see the shadows of broken oaths
\end{itemize}

\subsection{Witchcraft Follower Creation}

To create a Witchcraft follower:

\begin{enumerate}
    ℹtem Choose a follower type (Oathsworn, Threshold−Keeper, or Shadow−Speaking)
    ℹtem Determine their connection to the witch (how the pact was formed)
    ℹtem Give them a distinctive mark or trait related to their connection
    ℹtem Set their starting condition state (usually True−Bound or Clear−Sighted)
    ℹtem Determine one special action they can perform related to their connection
\end{enumerate}

\subsection{The Price of Power}

Witchcraft followers are not meant to be permanent. They are the living embodiment of the witch's choices, and like all things in the world of Witchcraft, they carry the weight of consequence. When a follower breaks, it is not just a loss−−−it is a story. When they thrive, they become the living proof that even in a world of shadows, there can be light.

In the end, the most powerful witchcraft is not in the spells cast or the pacts made, but in the bonds forged and the promises kept. For it is in these bonds that the true power of the Witchcraft system reveals itself−−−not as domination, but as the sacred responsibility of connection in a world that would rather forget.

\section*{The Unanswered Reflection}

When a follower dies while in service to a witch, the Unanswered Reflection intensifies. The witch must:

\begin{itemize}
    ℹtem Mark 3 Obligation segments immediately
    ℹtem Suffer −1 die to all rolls until they perform an act of penance for the fallen follower
    ℹtem Face the Shadow of the follower, who will appear at inopportune moments to remind them of their failure
\end{itemize}

The Shadow of the follower is not an enemy, but a manifestation of the Unanswered Reflection−−−the weight of a bond broken too soon. They do not attack, but they \textit{are} the reminder that every choice has consequences, and every bond carries a price.

\section*{The Living Pact}

Witchcraft followers are the living embodiment of the pact between witch and Patron. They are not tools to be used, but companions to be honored. When a witch respects this bond, the power of the pact grows stronger. When they take it for granted, the shadows deepen.

In the end, the true measure of a witch is not in the spells they cast, but in the bonds they keep. For in the world of Witchcraft, every promise is a thread, and the world itself is woven from the threads of unbroken oaths.

\vspace{1em}
\noindent\textit{"The follower is not your tool, but your mirror. In them, you see the weight of your choices. In their eyes, you see the cost of your power."} \\
−−− The Gallow's Bell, as recorded in the Book of the Unbroken Oath

% Appendix: The Unbroken Chord
% Requires: geometry, titlesec, enumitem, tcolorbox (skins), xcolor
% Compile with pdflatex (no fontspec)



\end{document}
