\section{Khemesh — The Abyssal Maw}

\subsection{Domain and Epithets}

Khemesh is not merely a lord of the depths but the hunger beneath them, a pressure older than seas. Those who bargain with him are marked by the abyss—seen in the way shadows cling, in the whispers heard when no voice speaks, in the certainty that all things will sink. He is the classic warlock's patron: distant, vast, and utterly indifferent to mortal morality. To serve Khemesh is to accept that the universe is cold, that the deep does not hate—it simply consumes.

In the trench without light, the Maw waits. Even silence drowns.

\subsection{Cultural Origins}

Khemesh appears wherever depth meets pressure and hunger:

\begin{itemize}
\item In \textbf{Zakov}, he is the Drowned Prince. Pirates toss a silver coin overboard before a raid—if it sinks, Khemesh approves.
\item In \textbf{Linn}, he is the Reef-Father. Fishermen leave the first catch of the day to him, cutting a piece of flesh and letting it sink.
\item In \textbf{Kahfagia}, he is the Pressure Below, invoked by divers and wreck-salvagers. His symbol is a black pearl.
\item In \textbf{Aeler}, he is the Deep Listener. Miners whisper to him before descending into new shafts. The stone answers—or does not.
\end{itemize}

\subsection{Warrior Archetype: The Deep Warrior}

A Runic Warrior of Khemesh is not a hunter of the surface. They are a \textit{Deep Warrior}—an eldritch knight who fights in crushing darkness, who carries the pressure of the abyss in their lungs and the silence of the trench in their blade. Their Codex is a shard of abyssal glass or a diving bell's valve. Their Thiasos is a drowned spirit that speaks in bubbles, a pressure-eel that senses weakness, or a sentient blade forged from the anchor of a sunken ship.

\subsection{Codex Form}

The Deep Warrior's Codex is born of depth:

\begin{itemize}
\item \textbf{Abyssal Glass:} A shard of black glass that shows no reflection. It is cold to the touch. It whispers when pressure changes.
\item \textbf{Diving Bell Valve:} A brass valve from a sunken diving bell. It turns to release pressure—or to seal it in.
\item \textbf{Black Pearl:} A pearl the color of a bruised storm. It is warm when the deep is near. It grows cold when danger passes.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Drowned Spirit:} The ghost of a sailor who died in darkness. It cannot speak, but it can point. It knows where the pressure is heaviest.
\item \textbf{Pressure-Eel:} A translucent eel that slides through cracks in stone. It senses weakness in structures and flesh alike. It is blind. It does not need to see.
\item \textbf{The Abyssal Glass (Sentient):} A shard of black glass that speaks in the voice of drowned captains. It can extinguish light in a Near zone once per session.
\end{itemize}

\subsection{Core Mechanic: The Crushing Depth}

When the Deep Warrior invokes a Rite, they may choose to draw on the pressure of the deep. For each segment of Obligation they mark beyond the Rite's cost, they gain +1 die to the roll (maximum +3). The pressure does not forgive. The pressure does not forget.

When the Deep Warrior is in darkness or underwater, they gain +1 die to all physical actions. The deep is their ally—or their grave.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Crushing Whisper (5 XP)}

\textbf{Tags:} [FEAR], [DEBUFF], [AREA], [SILENCE]

\textbf{Threshold:} A broken shell filled with ink-dark water.

Choose one: whisper of the trench (target hears impossible echoes and suffers -1 die on their next action) or crushing silence (establish an oppressive silence in a zone; enemies in the zone suffer -1 die to coordination or morale-driven actions).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Abyssal Grasp (7 XP)}

\textbf{Tags:} [BIND], [PRESSURE], [DEBUFF]

\textbf{Threshold:} A length of rusted chain submerged in water.

Target is pinned by invisible crushing force: treat as [ENTANGLE] with Great Effect if underwater or confined.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Alien Gaze (9 XP)}

\textbf{Tags:} [KNOWLEDGE], [TRUTH], [BURDEN], [FEAR]

\textbf{Threshold:} None—requires standing in absolute darkness or near deep water.

You perceive the world as Khemesh does—fractured, alien, crushing. Gain +2 dice to Notice and Arcana, and may ask one "true nature" question about a foe or structure. When the scene ends, you suffer Exposure +1 as your perception warps.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Pressure's Toll (8 XP)}

\textbf{Tags:} [TRANSFER], [DEBT], [DEBUFF]

\textbf{Threshold:} A functioning mechanism and a living creature's hair or blood.

Transfer inefficiency. You may take 1 Fatigue and reduce the DV of your next crafting or repair roll by 1, or touch an enemy to impose -1 die on their next physical action as their muscles "optimise" poorly.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Sundering Deep (14 XP)}

\textbf{Tags:} [PRESSURE], [AREA], [FATE], [DEATH]

\textbf{Threshold:} A sealed vessel of abyssal water, broken open.

Reality in the zone folds inward like the crushing deep: enemies act at Desperate Position by default. Each beat, the GM may force 1 SB (Spades/Clubs favoured) as the deep pressures reality. Structures, vessels, or wards fracture as if under immense weight.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Crushing Silence (6 XP)}

\textbf{Requirement:} Khemesh

Once per session as a reaction when a creature within Near range speaks or casts a spell, impose a bubble of oppressive silence around them for one exchange. They cannot speak, and any spell with a verbal component fails. Mark 1 Fatigue after using this talent.

\subsubsection{Deep Sight (4 XP)}

\textbf{Requirement:} Khemesh

You can see in darkness as if it were dim light. In absolute darkness, you are not blind—you perceive pressure and movement instead of light. This is a passive ability.

\subsubsection{Pressure Adaptation (8 XP)}

\textbf{Requirement:} Khemesh, Body 3+

You are immune to underwater combat penalties. You can hold your breath for twice as long as normal, and you ignore the first environmental penalty from pressure (diving, crushing depths, cave-ins).

\subsection{Roleplaying a Deep Warrior}

\begin{itemize}
\item You do not fear the dark. You were trained in it. The dark is where you hunt.
\item You may be called to salvage sunken wrecks, explore flooded ruins, or silence those who speak too loudly in quiet places.
\item You may struggle with the isolation. The deep is lonely. Khemesh does not offer companionship. He offers pressure.
\item You may be asked to let something sink—a memory, a hope, a friend. The deep takes. The deep does not return.
\end{itemize}

\subsection{Sample NPC: Captain Sefu of the Sunken Fleet}

\textbf{Archetype:} Deep Warrior

\textbf{Patron:} Khemesh

\textbf{Codex:} A shard of abyssal glass that shows no reflection

\textbf{Thiasos:} A drowned spirit that points toward pressure

\textbf{Personality:} Quiet, cold, inexorable. He has not seen the sun in seven years. The deep is his sky.

\textbf{Conflict:} A ship has sunk with a valuable cargo—and with living crew trapped in an air pocket. The cargo belongs to a patron who paid Sefu to retrieve it. The crew are strangers. The pressure is rising. The deep does not distinguish.

\subsection{GM Guidance: Intrusions for Khemesh}

\begin{itemize}
\item Water pressure quadruples in the room—wooden objects groan, glassware cracks, ears pop. The deep is close.
\item A nightmare tentacle manifests from a nearby body of water and steals a shiny object. It is not hostile. It is curious.
\item The party begins to sink—not physically, but emotionally. Each character marks 1 Fatigue or confesses a secret fear aloud. The deep listens.
\item The Deep Warrior's abyssal glass shows a face that is not theirs. It is smiling. It knows something they do not.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
