\section{Xhak'Thul — The Thunder-Speaker}

\subsection{Domain and Epithets}

Before words, there was thunder. Before scripture, lightning carved its commandments into the bones of the world. Xhak'Thul is the Voice That Splits Mountains, the Drum That Shakes Foundations, a power born from the first storm that ever raged across the primal earth. His faithful are whirlwinds of action—they do not study; they dance, they shout, they move with the sky's fury.

Xhak'Thul does not whisper. He roars. His gifts are not gentle miracles but raw terror, awe, and vitality ripped from nature's heart. To serve him is to become the storm—to move without hesitation, to strike without warning, to be the loudest voice in any room and to make that room tremble.

\subsection{Cultural Origins}

Xhak'Thul speaks wherever the sky meets the earth in violence:

\begin{itemize}
\item In the deep \textbf{Vhasian} highlands, he is the Storm-Father, invoked by herders who watch lightning split the sky above their flocks.
\item In \textbf{Ykrul}, he is the Thunder-Hoof, the sound of a thousand horses stampeding across the Violet Steppe.
\item In \textbf{Linn}, he is the Raging Tide, the storm that comes without warning and leaves only wreckage in its wake.
\item In \textbf{Xhak'Thul's own name}—spoken only in a whisper, never in a shout—is said to be the first sound the world ever made.
\end{itemize}

\subsection{Warrior Archetype: The Storm-Sworn}

A Runic Warrior of Xhak'Thul is not a scholar of weather patterns. They are a \textit{Storm-Sworn}—a primal warrior who carries the tempest in their blood, who shouts as a weapon, who moves with the sudden, violent grace of lightning. Their Codex is a drum made from storm-felled wood or a banner that twists in unfelt wind. Their Thiasos is a storm-spirit that rides the wind, a drum that beats itself, or a sentient blade forged from a lightning-struck spear.

\subsection{Codex Form}

The Storm-Sworn's Codex is an instrument of thunder:

\begin{itemize}
\item \textbf{Storm Drum:} A hand-drum made from the hide of a beast killed by lightning and the wood of a tree struck by the same bolt. It rumbles when storms approach.
\item \textbf{Thunder Banner:} A torn flag on a broken pole. It twists in wind that no one else can feel. When planted, it calls the storm.
\item \textbf{Lightning-Scarred Spear:} A spear whose wooden shaft is marked by a lightning strike. The mark glows faintly during storms.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Storm-Spirit:} A translucent figure of wind and static. It rides the air currents, can carry whispers across distances, and crackles when angered.
\item \textbf{Drum-Spirit:} A drum that beats itself—slowly when calm, rapidly when danger nears. It can be heard only by the Runekeeper and their allies.
\item \textbf{The Thunder Drum (Sentient):} A drum that speaks in drumbeats—slow for warning, rapid for danger, a single, sharp crack for "strike now."
\end{itemize}

\subsection{Core Mechanic: The Storm's Fury}

When the Storm-Sworn takes the initiative in combat (acts first in a scene), they gain +1 die to their first action. The storm does not wait.

When the Storm-Sworn is in an active storm (rain, wind, thunder, lightning), they gain +1 die to all physical actions. The storm knows its own.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Thunderous Presence (4 XP)}

\textbf{Tags:} [FEAR], [COMMAND], [AREA], [PRESSURE]

\textbf{Threshold:} A drum made from storm-felled wood or any suitable improvised percussion.

Your presence becomes overwhelming. All NPCs within Near range must test Resolve (DV 2) or suffer -1 die to social actions due to awe or terror for the scene. Allies gain +1 die to Intimidation or Command actions.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Storm's Step (5 XP)}

\textbf{Tags:} [SPEED], [FEAR], [BOOST]

\textbf{Threshold:} Earth from a lightning strike site, worn boot leather.

You move with the storm's speed. Gain +2 dice to Athletics or Stealth rolls. The first attack you make this scene gains +1 Effect.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shattered Sky (8 XP)}

\textbf{Tags:} [AREA], [FORCE], [FEAR], [WEATHER]

\textbf{Threshold:} A shard of meteoric iron or metal touched by a severe storm.

You call a localized, supernatural storm in a zone. Choose one effect: lightning strike (single target, Harm 2), wind blast (knockback, disperses arrows or scatters foes), or hail barrage (Harm 1, AREA). All enemies in the zone suffer -1 die to actions from chaos and distraction.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Fear-Eater (9 XP)}

\textbf{Tags:} [FEAR], [HEAL], [COMMAND], [BOOST]

\textbf{Threshold:} Soil or dust from a place of great terror (battlefield, prison, ruins).

You feed on the fear around you. For each Frightened target in the scene, gain +1 die on your next action (max +3). If you successfully Intimidate a target during this scene, you heal 1 Fatigue.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the World's Drum (14 XP)}

\textbf{Tags:} [AREA], [FEAR], [COMMAND], [TRANSFORM], [WEATHER]

\textbf{Threshold:} A drum fashioned from lightning-struck wood, with a hide from a beast killed by natural force.

You become an incarnation of the storm for one scene: gain +2 dice to rolls based on strength, speed, and intimidation; your anger echoes as thunder (attacks against you may be answered by lightning, GM discretion, usually 1 SB and Harm 1 to the attacker); enemies test Resolve at -2 dice to resist Fear effects you create. Begin a 6-segment Natural Disasters clock.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Storm Caller (6 XP)}

\textbf{Requirement:} Xhak'Thul

Once per scene, as a free action, you may cause a peal of thunder to roll across the zone. All enemies within Near must test Resolve (DV 3) or lose their next minor action.

\subsubsection{Lightning Reflexes (4 XP)}

\textbf{Requirement:} Xhak'Thul, Body 3+

You gain +1 die to defense rolls against ranged attacks (arrows, bolts, thrown weapons). You see the path of the projectile before it flies.

\subsubsection{Fear Eater (8 XP)}

\textbf{Requirement:} Xhak'Thul, Presence 3+

When you successfully Intimidate a target, you gain 1 Boon (once per scene). The fear is not just a weapon—it is sustenance.

\subsection{Roleplaying a Storm-Sworn}

\begin{itemize}
\item You do not wait. You act. Hesitation is a lie the wind tells to the still.
\item You may be called to break sieges, scatter formations, or terrify enemies into surrender. Your voice is a weapon. Your presence is a warning.
\item You may struggle with the aftermath. The storm passes. You remain. The silence after thunder is heavy.
\item You may be asked to be gentle—to speak softly, to wait, to hold back. The storm does not understand gentleness. You must learn.
\end{itemize}

\subsection{Sample NPC: Storm-Speaker Urum of the Violet Steppe}

\textbf{Archetype:} Storm-Sworn

\textbf{Patron:} Xhak'Thul

\textbf{Codex:} A drum made from a tree struck by lightning

\textbf{Thiasos:} A storm-spirit that rides the wind ahead of him

\textbf{Personality:} Loud, restless, generous. He has never whispered a word in his life. He does not intend to start.

\textbf{Conflict:} A drought has gripped the steppe. The clans are dying. Urum can call rain—but the storm will not be gentle. It will flood. It will kill. The price of mercy is destruction. The storm does not negotiate.

\subsection{GM Guidance: Intrusions for Xhak'Thul}

\begin{itemize}
\item A sudden storm appears on a clear day. The wind shifts direction three times in an hour. Xhak'Thul is watching.
\item The Storm-Sworn's drum beats once—loud, sharp, unexpected. Something has noticed them. Something hungry.
\item Lightning strikes near the party—not aimed, not random. A warning. The sky is listening.
\item The Storm-Sworn's voice cracks. Not from strain—from pressure. The storm is inside them now. It wants out.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
