\section{Raéyn — Mistress of the Sea}

\subsection{Domain and Epithets}

Raéyn is the tempestuous goddess of the sea, the restless tide that carries news between shores and the promise of change between lives. She is mother to all who sail, her voice the wind that fills sails and her moods the storms that test every mariner's resolve. Yet Raéyn's heart is torn by her greatest tragedy: her son Khemesh, the Kraken of the Depths, who embodies the crushing inevitability of the ocean's dark heart. Where Raéyn brings change and opportunity, Khemesh brings the final, inescapable pressure that grinds all things to nothing.

Sailors pray to Raéyn for safe passage and favourable winds, but whisper Khemesh's name when seeking to lay the dead to rest beneath the waves. She is passionate, mercurial, and fiercely protective of those who respect her domain. Cross her, and the sea itself becomes your enemy.

\subsection{Cultural Origins}

Raéyn is honored wherever salt water meets shore:

\begin{itemize}
\item In \textbf{Kahfagia}, she is the Lantern-Tide. Pilots light a beacon to Raéyn before every night voyage. Her symbol is a wave curling into a spiral.
\item In \textbf{Linn}, she is the Whale-Mother. Funerals at sea involve a song and a single white stone cast into the deep.
\item In \textbf{Zakov}, she is the Salt-Queen. Pirates carve her face on their figureheads—one eye open (fair seas), one eye closed (storms ahead).
\item In \textbf{Sidhi} ports, she is the Tide-Father's Sister, invoked before any voyage that crosses the open ocean.
\end{itemize}

\subsection{Warrior Archetype: The Tide-Knight}

A Runic Warrior of Raéyn is not a storm-caller in the manner of Xhak'Thul. They are a \textit{Tide-Knight}—a warrior who fights on rolling decks and slippery shores, who reads the currents of battle as a pilot reads the sea, who falls only to rise with the next wave. Their Codex is a ship's wheel or an anchor. Their Thiasos is a selkie-spirit, a seabird that carries warnings, or a sentient blade forged from a ship's salvaged prow.

\subsection{Codex Form}

The Tide-Knight's Codex is born of the sea:

\begin{itemize}
\item \textbf{Ship's Wheel:} A small wooden wheel from a vessel that survived a storm. It turns smoothly. It never sticks.
\item \textbf{Anchor Charm:} A small iron anchor on a leather cord. It is heavy. It grounds the wearer against being thrown.
\item \textbf{Salt-Crusted Compass:} A brass compass filled with seawater instead of oil. The needle spins—but it always points to the nearest safe harbor.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Selkie-Spirit:} A seal that can shed its skin to become a woman with hair like kelp. She knows the currents. She knows the hidden reefs.
\item \textbf{Seabird Spirit:} A grey gull or tern that flies ahead of the ship. It returns with warnings—storms, pirates, safe passages.
\item \textbf{The Salt Compass (Sentient):} A compass that speaks in the rustle of waves. It can find the nearest safe harbor once per journey, but it will ask for a story of the sea in return.
\end{itemize}

\subsection{Core Mechanic: The Rising Tide}

When the Tide-Knight is on or near a body of water (river, lake, sea), they gain +1 die to all physical actions. The water supports them.

When the Tide-Knight is knocked prone or loses their footing, they may spend 1 Boon to immediately rise as a free action. The wave that knocked them down also lifts them up.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Wave-Tide (5 XP)}

\textbf{Tags:} [MOVEMENT], [NAVIGATION], [BOOST], [RISK]

\textbf{Threshold:} A knotted length of salt-twine brushed with seawater, or a shell held to the ear while facing the wind.

Choose one: tidemark's blessing (treat slick, swaying, or water-slicked footing as stable; gain +1 die on boarding, balance, or shipboard movement) or whispering currents (learn the safest near-term route across water or coastline, or gain +1 die to navigation checks for this scene).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Tide's Favor (4 XP)}

\textbf{Tags:} [MOVEMENT], [BALANCE], [BOOST]

\textbf{Threshold:} A shell or stone taken from the high-tide line.

You feel the rhythm of the water. For one scene, ignore one level of environmental penalty from moving water (choppy seas, strong currents, slippery decks). You cannot be knocked prone by water alone.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Changing Tide (8 XP)}

\textbf{Tags:} [TERRAIN], [AREA], [MOVEMENT], [PRESSURE]

\textbf{Threshold:} A handful of pebbles cast in a crescent.

Bias currents and water levels in the zone. Those moving with the tide gain +1 die; those moving against suffer -1 die. Small craft must test to hold position. Create a 6-segment Tidal Influence clock.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Wave-Road Consecration (7 XP)}

\textbf{Tags:} [GUIDANCE], [BOOST], [BOND], [FATE]

\textbf{Threshold:} Two sea-glass markers dropped overboard at start and end.

Consecrate a wave-road between two visible points. Allies gain +2 dice on travel, evade, or carry actions at sea. Designated pursuers suffer -1 die to intercept. One active wave-road at a time. Create an 8-segment Blessed Passage clock.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Storm-Queen's Hand (13 XP)}

\textbf{Tags:} [WEATHER], [AREA], [COMMAND], [FEAR], [BOOST]

\textbf{Threshold:} A vial of rainwater gathered at three crossings.

Shape a storm-band over the zone. Choose two modes at cast; switch one once per scene: propulsion (vessel gains +1 band of movement per beat), concealment (veil of rain/spray; ranged targeting impaired; -1 die to hostile sighting), or smite (once per beat, lash with wave or lightning as [AREA] hazard).

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Tide-Step (4 XP)}

\textbf{Requirement:} Raéyn

Once per session as a free action, you may treat any water-based difficult terrain as stable ground for one exchange. Alternatively, you may move across the surface of water as if it were solid for up to Near distance. Mark 1 Fatigue after using this talent.

\subsubsection{Storm Glass (6 XP)}

\textbf{Requirement:} Raéyn

You can predict the weather for the next hour with perfect accuracy. You know when a storm is coming, how long it will last, and roughly how severe it will be. This is a passive ability.

\subsubsection{Kraken's Warning (8 XP)}

\textbf{Requirement:} Raéyn, Spirit 3+

Once per session, when a sea creature or Khemesh's influence would ambush you, you sense it one round in advance. Your party cannot be surprised. The warning is a chill down the spine—nothing more, nothing less.

\subsection{Roleplaying a Tide-Knight}

\begin{itemize}
\item You do not fight against the current. You read it. You use it. The sea is not your enemy—it is your context.
\item You may be called to escort ships, guide refugees across rivers, or fight on the decks of vessels in stormy weather.
\item You may struggle with the weight of Raéyn's grief. Her son is the deep. The sea is both mother and grave.
\item You may be asked to choose—between tide and trench, between change and pressure, between Raéyn and Khemesh. The sea does not forgive a divided heart.
\end{itemize}

\subsection{Sample NPC: Captain Mira of the Salt Compass}

\textbf{Archetype:} Tide-Knight

\textbf{Patron:} Raéyn

\textbf{Codex:} A salt-crusted compass that points to safe harbor

\textbf{Thiasos:} A seabird spirit that warns of storms

\textbf{Personality:} Weather-worn, laughing, fierce. She has lost three ships to storms. She has never lost a crew.

\textbf{Conflict:} A storm is coming—not a natural storm, but a summoned one. Khemesh is waking. Raéyn wants Mira to calm the waters. Khemesh wants Mira to dive. The sea is her mother. The deep is her brother. She cannot choose.

\subsection{GM Guidance: Intrusions for Raéyn}

\begin{itemize}
\item A sudden fog rolls in—all navigation rolls at -2 dice for the scene. Raéyn is hiding something.
\item The tide turns early, stranding a vessel you need. The sea is impatient.
\item A dolphin, seal, or seabird follows the party for a day. It will deliver one short message to the nearest follower of Raéyn if fed something shiny.
\item The Tide-Knight's compass spins without stopping. The nearest safe harbor is not safe—not anymore. Something has changed.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
