\section{Rites of the Deep Patrons}

The Deep patrons—Khemesh, Xhak'Thul, and Raéyn—offer Rites that draw on the crushing pressure of the abyss, the wild fury of the storm, and the relentless rhythm of the tide. These Rites are elemental, overwhelming, and often dangerous to the caster as well as the target.

\subsection{Khemesh — Deep Warrior Rites}

\subsubsection{Low Rite: Rite of the Crushing Whisper (5 XP)}

\textbf{Tags:} [FEAR], [DEBUFF], [AREA], [SILENCE]

\textbf{Threshold:} A broken shell filled with ink-dark water.

Choose one:
\begin{itemize}
\item \textit{Whisper of the Trench:} Target hears impossible echoes and suffers -1 die on their next action.
\item \textit{Crushing Silence:} Establish an oppressive silence in a zone; enemies in the zone suffer -1 die to coordination or morale-driven actions.
\end{itemize}

\textbf{Push It:} (Whisper) Echoes coil in your own skull—take Fatigue 1, but the target also loses their next minor action. (Silence) A single enemy's voice is stolen entirely for the scene.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Abyssal Grasp (7 XP)}

\textbf{Tags:} [BIND], [PRESSURE], [DEBUFF]

\textbf{Threshold:} A length of rusted chain submerged in water.

Target is pinned by invisible crushing force: treat as [ENTANGLE] with Great Effect if underwater or confined.

\textbf{Push It:} Inflict Fatigue 1 on the target in addition to the restraint.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Alien Gaze (9 XP)}

\textbf{Tags:} [KNOWLEDGE], [TRUTH], [BURDEN], [FEAR]

\textbf{Threshold:} None—requires standing in absolute darkness or near deep water.

You perceive the world as Khemesh does—fractured, alien, crushing. Gain +2 dice to Notice and Arcana, and may ask one "true nature" question about a foe or structure (e.g., "What is this creature's vulnerability?" or "What holds this wall together?").

\textbf{Cost:} When the scene ends, you suffer Exposure +1 as your perception warps.

\textbf{Push It:} Extend the vision to one ally, but both take Fatigue 1.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Pressure's Toll (8 XP)}

\textbf{Tags:} [TRANSFER], [DEBT], [DEBUFF]

\textbf{Threshold:} A functioning mechanism and a living creature's hair or blood.

Transfer inefficiency. Choose one:
\begin{itemize}
\item Take 1 Fatigue and reduce the DV of your next crafting or repair roll by 1.
\item Touch an enemy to impose -1 die on their next physical action as their muscles "optimise" poorly.
\end{itemize}

\textbf{Push It:} The transfer becomes permanent: you lose 1 Fatigue capacity (reduce Body by 1 for Fatigue track only) and the target gains a minor compulsive behaviour (GM's choice). Mark +2 Obligation.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Sundering Deep (14 XP)}

\textbf{Tags:} [PRESSURE], [AREA], [FATE], [DEATH]

\textbf{Threshold:} A sealed vessel of abyssal water, broken open.

Reality in the zone folds inward like the crushing deep:
\begin{itemize}
\item Enemies act at Desperate Position by default.
\item Each beat, the GM may force 1 SB (Spades/Clubs favoured) as the deep pressures reality.
\item Structures, vessels, or wards fracture as if under immense weight.
\end{itemize}

\textbf{Push It:} For one beat, declare a single enemy "crushed" (severe harm/effect). You immediately suffer Fatigue 2 and +1 Obligation.

\textbf{Obligation:} 3

\subsection{Xhak'Thul — Storm-Sworn Rites}

\subsubsection{Low Rite: Rite of the Thunderous Presence (4 XP)}

\textbf{Tags:} [FEAR], [COMMAND], [AREA], [PRESSURE]

\textbf{Threshold:} A drum made from storm-felled wood or any suitable improvised percussion.

Your presence becomes overwhelming. All NPCs within Near range must test Resolve (DV 2) or suffer -1 die to social actions due to awe or terror for the scene. Allies gain +1 die to Intimidation or Command actions.

\textbf{Push It:} Extend the effect to Far range, but you mark 1 Fatigue and mark 1 SB (Hearts) as your presence becomes difficult to control.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Storm's Step (5 XP)}

\textbf{Tags:} [SPEED], [FEAR], [BOOST]

\textbf{Threshold:} Earth from a lightning strike site, worn boot leather.

You move with the storm's speed. Gain +2 dice to Athletics or Stealth rolls. The first attack you make this scene gains +1 Effect.

\textbf{Push It:} Gain +3 dice and move as a blur, becoming difficult to track or clearly remember; mark 1 SB (Diamonds).

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shattered Sky (8 XP)}

\textbf{Tags:} [AREA], [FORCE], [FEAR], [WEATHER]

\textbf{Threshold:} A shard of meteoric iron or metal touched by a severe storm.

You call a localised, supernatural storm in a zone. Choose one effect:
\begin{itemize}
\item \textit{Lightning Strike:} Single target, Harm 2.
\item \textit{Wind Blast:} Knockback, dispersing arrows or scattering foes.
\item \textit{Hail Barrage:} Harm 1 (AREA).
\end{itemize}
All enemies in the zone suffer -1 die to actions from chaos and distraction.

\textbf{Push It:} Double the chosen effect's scale or potency (e.g., Harm 3, large knockback), but the effect also applies to allies in the zone; mark 2 SB (Clubs).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Fear-Eater (9 XP)}

\textbf{Tags:} [FEAR], [HEAL], [COMMAND], [BOOST]

\textbf{Threshold:} Soil or dust from a place of great terror (battlefield, prison, ruins).

You feed on the fear around you. For each Frightened target in the scene, gain +1 die on your next action (max +3). If you successfully Intimidate a target during this scene, you heal 1 Fatigue.

\textbf{Push It:} Absorb all palpable fear in the scene, gaining a flat +3 dice bonus to your next action, but mark 1 permanent Scar related to fear. Mark 1 SB (Hearts).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the World's Drum (14 XP)}

\textbf{Tags:} [AREA], [FEAR], [COMMAND], [TRANSFORM], [WEATHER]

\textbf{Threshold:} A drum fashioned from lightning-struck wood, with a hide from a beast killed by natural force.

You become an incarnation of the storm for one scene:
\begin{itemize}
\item Gain +2 dice to rolls based on strength, speed, and intimidation.
\item Your anger echoes as thunder; attacks against you may be answered by lightning (GM discretion, usually 1 SB and Harm 1 to the attacker).
\item Enemies must test Resolve at -2 dice to resist Fear effects you create.
\end{itemize}
Start a 6-segment Natural Disasters clock.

\textbf{Push It:} The transformation lingers after the scene, and your emotional state causes minor environmental effects (windows rattle, small fires flicker). Mark 2 SB (Diamonds) and advance the clock by 2 segments.

\textbf{Obligation:} 3

\subsection{Raéyn — Tide-Knight Rites}

\subsubsection{Low Rite: Rite of the Wave-Tide (5 XP)}

\textbf{Tags:} [MOVEMENT], [NAVIGATION], [BOOST], [RISK]

\textbf{Threshold:} A knotted length of salt-twine brushed with seawater, or a shell held to the ear while facing the wind.

Choose one:
\begin{itemize}
\item \textit{Tidemark's Blessing:} Treat slick, swaying, or water-slicked footing as stable for this scene. Gain +1 die on boarding, balance, or shipboard movement. Create a 4-segment Tide's Favor clock to ignore one level of difficult terrain.
\item \textit{Whispering Currents:} Learn the safest near-term route across water or coastline, or gain +1 die to navigation checks for this scene. If Khemesh's influence is present, suffer -1 die from conflicting currents.
\end{itemize}

\textbf{Push It:} (Tidemark) Extend to one ally in Close for one beat, but generate 1 SB (Spades) as the deck shifts. (Currents) Also learn the fastest route, but mark Exposure +1 (leaving a telltale wake).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Tide's Favor (4 XP)}

\textbf{Tags:} [MOVEMENT], [BALANCE], [BOOST]

\textbf{Threshold:} A shell or stone taken from the high-tide line.

You feel the rhythm of the water. For one scene, ignore one level of environmental penalty from moving water (choppy seas, strong currents, slippery decks). You cannot be knocked prone by water alone.

\textbf{Push It:} Extend the effect to one ally within Close range, but mark 1 SB (Hearts) as the tide's favor becomes known to others.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Changing Tide (8 XP)}

\textbf{Tags:} [TERRAIN], [AREA], [MOVEMENT], [PRESSURE]

\textbf{Threshold:} A handful of pebbles cast in a crescent.

Bias currents and water levels in the zone. Those moving with the tide gain +1 die; those moving against suffer -1 die. Small craft must test to hold position. Create a 6-segment Tidal Influence clock.

\textbf{Push It:} Brief surge or drawdown (one beat): open a ford or swamp a skiff; mark +1 Obligation.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Wave-Road Consecration (7 XP)}

\textbf{Tags:} [GUIDANCE], [BOOST], [BOND], [FATE]

\textbf{Threshold:} Two sea-glass markers dropped overboard at start and end.

Consecrate a wave-road between two visible points. Allies gain +2 dice on travel, evade, or carry actions at sea. Designated pursuers suffer -1 die to intercept. One active wave-road at a time. Create an 8-segment Blessed Passage clock.

\textbf{Push It:} Extend the route's favour to an adjacent leg for one beat; mark +1 Obligation.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Storm-Queen's Hand (13 XP)}

\textbf{Tags:} [WEATHER], [AREA], [COMMAND], [FEAR], [BOOST]

\textbf{Threshold:} A vial of rainwater gathered at three crossings.

Shape a storm-band over the zone. Choose two modes at cast; switch one once per scene:
\begin{itemize}
\item \textit{Propulsion:} Vessel gains +1 band of movement per beat (or +1 Effect to manoeuvres).
\item \textit{Concealment:} Veil of rain and spray; ranged targeting impaired; -1 die to hostile sighting.
\item \textit{Smite:} Once per beat, lash with wave or lightning as [AREA] hazard.
\end{itemize}

\textbf{Push It:} Add a third mode for one beat, then the GM spends 1 SB on collateral; mark +1 Obligation.

\textbf{Obligation:} 3

\subsection{Rite Synergies}

Some Rites work well together for Deep patrons:

\begin{itemize}
\item \textbf{Rite of the Crushing Whisper + Rite of the Abyssal Grasp:} Silence a foe, then bind them in crushing pressure. They cannot call for help. They cannot move. The deep takes them.
\item \textbf{Rite of the Shattered Sky + Rite of the Fear-Eater:} Call a storm, then feed on the terror it creates. The lightning is your harvest; the thunder is your table.
\item \textbf{Rite of the Changing Tide + Rite of the Wave-Road Consecration:} Shift the currents, then consecrate the path. Your allies move with the water; your enemies struggle against it.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
