% chapter/appendix.tex
% Appendix -- Quick Reference, Printable Tools, and Sealed Secrets
% Rewritten in the voice of Lerris Fenwood, with marginalia from the Gray Wanderer,
% Saikou Ira, and Dusana of the Raven Road.

\epigraph{``An appendix is not a graveyard for unused ideas. It is a toolkit for the desperate. Keep it close. Keep it honest. Keep it sealed when it needs to be.''}{---Lerris Fenwood, from the front matter of his own ledger}

What follows is not lore. It is \textbf{utility}. Tables you can glance at during play. Cards you can print and cut. A sealed section for the GM's eyes only. And a final essay from me, Lerris Fenwood, on the counting of wounds -- because I have counted more than I care to admit.

Use these pages when the players have taken a road you did not map. They will. They always do.

\section*{Appendix A -- Quick Reference Tables}

\subsection*{Threat Level Summary}

\noindent (Repeated from Chapter 7 for ease of use.)

\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Threat Level} & \textbf{Typical Use} & \textbf{Base Harm} & \textbf{Base Fatigue} & \textbf{Attribute Adj.} \\
\hline
0 & Non-combat, flavor & 1--2 & 1 & --1 \\
1 & Standard minion & 3 & 2 & 0 \\
2 & Elite, lieutenant & 4--5 & 3 & +1 \\
3 & Mini-boss & 6--7 & 4 & +2 \\
4+ & Major boss & 8+ & 5+ & +3 \\
\hline
\end{tabular}

\subsection*{Common Timer Sizes and Their Meaning}

\begin{tabular}{|l|l|}
\hline
\textbf{Segments} & \textbf{Typical Use} \\
\hline
4 & Short scene, minor threat, a single chase \\
6 & Standard encounter, a duel, a negotiation \\
8 & Extended fight, a dungeon level, a complex ritual \\
10 & Epic boss, a siege, the finale of a campaign \\
\hline
\end{tabular}

\noindent\textbf{Rule of Thumb}: A 4-segment timer takes about 15--20 minutes of play. An 8-segment timer takes 40--60 minutes. Adjust for table pace.

\subsection*{Harm and Fatigue at a Glance}

\begin{tabular}{|l|l|}
\hline
\textbf{Harm Level} & \textbf{Effect} \\
\hline
1 & --1 die on related actions \\
2 & --1 die on most actions until treated \\
3 & Incapacitated or dying \\
\hline
\end{tabular}

\noindent\textbf{Fatigue}: Each point worsens Position by one step (Dominant → Controlled → Desperate). If already Desperate, each Fatigue instead imposes --1 die. Full Fatigue track (Body rating) converts to Harm +1 and clears all Fatigue.

\subsection*{Armor Conversion (Quick)}

\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Armor} & \textbf{Harm 1 →} & \textbf{Harm 2 →} & \textbf{Harm 3 →} \\
\hline
Light & Fatigue 1 & Harm 1 + Fatigue 1 & Harm 2 + Fatigue 1 \\
Medium & Fatigue 1 & Fatigue 1 & Harm 2 + Fatigue 1 \\
Heavy & Fatigue 1 & Fatigue 1 & Fatigue 2 \\
\hline
\end{tabular}

\gray{I have seen a player forget the armor conversion three times in one fight. The fourth time, I wrote it on an index card. The card saved his character's life. Do not be clever. Be clear.}

\section*{Appendix B -- Printable Adversary Card Template}

\noindent Print this page at 100\% scale. Cut along the hairlines. Fold and glue or laminate for durability.

\vspace{0.5cm}

\begin{center}
\fbox{\parbox{0.8\textwidth}{\centering
\textbf{Adversary Card -- Front}\\
\vspace{0.3cm}
\texttt{[Name]}\\
\texttt{[Threat Level] [Creature Type]}\\
\texttt{Harm: ] [ Fatigue: ]}\\
\texttt{Traits: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}\\
\texttt{Complications: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}\\
\texttt{Clock: \_\_\_\_\_\_\_\_ (segments) -- Tick on \_\_\_\_\_\_\_\_\_}
}}
\end{center}

\vspace{0.5cm}

\begin{center}
\fbox{\parbox{0.8\textwidth}{\centering
\textbf{Adversary Card -- Back}\\
\vspace{0.3cm}
\texttt{Weakness: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}\\
\texttt{Resolution (if not combat): \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}\\
\texttt{Drop: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}\\
\texttt{Marginal note: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_}
}}
\end{center}

\noindent\textbf{How to use}: Fill in the blanks before the session or during play. Keep a stack of blank cards behind the screen. I use a soft pencil; the Gray Wanderer uses blood (do not copy him).

\saikou{I prefer a silver mirror. The card is a backup. The mirror never lies.}

\section*{Appendix C -- Icon Legend}

\noindent Some creatures in this compendium use icon codes for quick reference. The following legend applies.

\begin{tabular}{|l|l|}
\hline
\textbf{Icon} & \textbf{Meaning} \\
\hline
[W] & Weakness -- silver, fire, a true name, etc. \\
[C] & Timer -- size and trigger \\
[L] & Lair -- special environmental action \\
[P] & Phase -- boss phase transition \\
[D] & Drop -- treasure or boon \\
[WIT] & Requires a witness to resolve \\
[NIN] & Involves the ninth -- counting, omission, taboo \\
\hline
\end{tabular}

\noindent Example: \emph{Slasher [W: silver] [C: Hunger 6 -- ticks on kill]}. Means the slasher is vulnerable to silver and has a Hunger timer that ticks each time it kills.

\lerris{I invented this legend after a player asked me the same question four times in one fight. The legend did not help. The fight did not end well. But the legend was not at fault.}

\section*{Appendix D -- Sealed Section (GM Eyes Only)}

\noindent The following pages are sealed. Open only when you need a secret adversary, a hidden weakness, or a final twist. The seal is not magical. It is a promise.

\subsection*{D.1 -- The Unnamed Adversary: The Ledger Lich (Tier IV)}

\noindent\textbf{Description}: A curator bound by indices; hands of ink, breath of dust. It appears in archives, counting-houses, and the little rooms where ledgers are kept. It does not fight. It \textbf{audits}.

\noindent\textbf{Stats}: Body 2, Wits 5, Spirit 5, Presence 4. Harm 6. Fatigue Spirit.

\noindent\textbf{Phase Timer }: [8] -- Citation → Injunction → Foreclosure → Redaction

\noindent\textbf{Moves}: \emph{Name Forfeiture} (steal a declared tag until dawn), \emph{Writ of Seizure} (lock a Diamond or gear under spectral seal), \emph{Errata} (retroactively worsen one PC's prior roll Outcome).

\noindent\textbf{Lair Actions}: \emph{Quiet Hours} (casting/rites generate +1 SB or fail softly), \emph{Reference Only} (movement through stacks costs an action unless proper tokens are presented), \emph{Cross-reference Collapse} (bookshelves reconfigure; split the party).

\noindent\textbf{Counters}: True names carved in quartz, burning a citation (sacrifice an index -- a map, a ledger, a vow) to cancel one Move.

\noindent\textbf{Drop}: A Key of Echoes (unlocks any door, but the door remembers) or a Tome of Convenient Knowledge (once per session, answer a single question about a creature's weakness).

\subsection*{D.2 -- The Unwritten Puzzle: The Ninth Name}

\noindent\textbf{Setup}: The party finds a sealed envelope with the words ``Open when the ninth bell rings'' written in a familiar hand (a dead mentor, a lost sibling). Inside is a single name.

\noindent\textbf{The Puzzle}: The name is not a name. It is a \textbf{count}. Each letter corresponds to a number (A=1, B=2, etc.). Sum the numbers. Divide by eight. The remainder is the number of the threshold where the envelope was sealed.

\noindent\textbf{The Secret}: The threshold is a door that has been bricked over in the Wardens`` archive. Behind it is a confession -- the mentor's final witness. The party must decide whether to read it aloud.

\noindent\textbf{Sealed Solution}: The name is \texttt{ALAYSE} (A=1, L=12, A=1, Y=25, S=19, E=5 → sum = 63. 63 ÷ 8 = 7 remainder 7. The threshold is the seventh door in the east corridor. The bricked-over door. Behind it, the confession: \emph{''I lied about the ninth step. It is not a door. It is a witness. The witness is you.``}

\saikou{I helped seal this page. The ink is still warm. The secret is not a secret. It is a debt. Pay it when you are ready.}

\section*{Appendix E -- On the Counting of Wounds (An Essay by Lerris Fenwood)}

\noindent I have killed seventeen creatures catalogued in this book. I have been wounded by twenty-three. The wounds heal. The counting does not.

When I was young, my father taught me to count my breaths. ''Eight,`` he said. ''Never nine. The ninth breath belongs to the Hollow.`` I did not believe him. Then I saw a man take the ninth breath in an Aeler tunnel. The air turned cold. The man's shadow stood up and walked away. The man did not follow. He simply sat down and waited. He is still waiting.

Adversaries are not bags of harm. They are \textbf{questions}. A ghoul asks: ''What are you willing to eat?`` A vampire asks: ''What are you willing to lose?`` A Oath-Keeper asks: ''What do you owe?`` The dice do not answer these questions. The players do. The GM listens. The ledger is written in the space between.

I keep a second ledger for the wounds that did not heal. The first entry is a scar on my left hand -- a ghoul's claw. The second entry is a silence I cannot name -- the debt I owe to the String-Mother. The third entry is a door I did not open. I write this now so that you, reader, will know that the compendium is not complete. The compendium is a witness. The wounds are the testimony.

Count to eight. Write what you can prove. Leave the ninth line empty. The Hollow will fill it when it is ready.

\lerris{This essay is not advice. It is a confession. Use it as you will.}

\section*{Appendix F -- Colophon}

\noindent This book was written on paper that had been a cargo manifest, a love letter, and a bandage. The ink is lamp-black cut with salt water and the ash of a burnt confession. The binding is gray cord tied with nine knots. The ninth knot is not cut.

The Gray Wanderer contributed his usual cynicism and a recipe for tea that keeps the Hollow from your doorstep (mostly). Saikou Ira insisted on keeping a silver mirror beside the writing desk; it fogged exactly twice, and he will not say what he saw. Dusana sang a lark into the rafters; the bird stayed for three days and left a single feather, which I have pressed between these pages.

As for me, Lerris Fenwood, I have seen the faces of the Unmasked. I have read the true ledger of the Congregation. I have learned that the ninth name is never spoken because it belongs to the one who will speak it last. When that day comes, I hope this book serves as witness.

Until then, the road is south. The ink is fresh. The lamp is borrowed.

\gray{You forgot the bread. Always leave bread for the threshold.}
\saikou{And a coin for the ferryman, just in case.}
\dusana{And a story for the fire. The rest will take care of itself.}
\lerris{Paid in full.}

\cleardoublepage


\chapter{Rite of the Echoing Vessel}

\begin{center}
\emph{The ink is silver. The mirror is cold. The debt is not mine.}
\end{center}

\vspace{1em}

\section*{Rite of the Echoing Vessel}
\emph{Varnek Karn, High Rite}

\begin{ritualbox}
\textbf{Components:} A vessel that has touched the dead---a lock of hair, a burial coin, a piece of jewelry worn at the moment of passing. A bowl of still water drawn from the river where they died, or from a threshold between waters (a confluence, a ford, a place where fresh meets salt). A gray thread, spun from wool that has been neither washed nor dyed, knotted nine times. A mirror that does not lie---silvered glass, still water, or the polished blade of a knife that has never drawn blood.

\medskip
\textbf{Obligation:} 3 segments. If the vision shows a debt unpaid, the caster must accept a Promise Timer [4] to settle it, or mark 1 Corruption. The Symbol of Varnek Karn grows cold during the rite; if it becomes warm, the dead are near, and the caster should not look away.
\end{ritualbox}

\vspace{1em}

You prepare the space at twilight, when the boundary between worlds is thinnest. You face east, toward the river, though you may not be near water. The dead do not need geography. They need direction.

You place the bowl on bare earth. If you are indoors, you set it on a stone that has never been worked by a tool. You fill it with water drawn from the place of death, or from the nearest running water you could reach. If you have neither, you use rainwater collected in a vessel that has never held wine. The dead are particular about vessels. They remember what has been drunk from them.

You take the gray thread. You knot it once, speaking the name of a dead you have known. You knot it twice, speaking the name of a dead you have forgotten. You knot it thrice, speaking the name of a dead you never met. You knot it four times, five times, six times, seven times, eight times---each knot a breath held, a word withheld, a question you do not ask.

The ninth knot you do not tie. You lay the thread across the rim of the bowl, one end in the water, the other in your hand. The ninth knot is the question. The dead will tie it, or they will not.

You place the vessel---the lock of hair, the burial coin, the locket---into the water. You do not speak its story. The dead already know. The dead were there.

You wait.

The water stills. The mirror you have placed beside the bowl begins to fog, not from warmth but from breath---breath that is not yours, breath that has not touched air in years. You do not wipe the mirror. You do not look away.

You see.

\vspace{1em}

\begin{visionbox}
\subsection*{Sunset Over the Black River}
\emph{What the confession did not say.}

\medskip
\noindent The voice came to Bryn in dreams---not as a memory, but as a presence. Malachai's honeyed rasp, sweet as rot, patient as a sealed tomb. She had heard it only once, in the vaulted chamber beneath Zahk''Kozahn, when she was small enough to fit in her mother's arms. She had not understood the words then. She understood them now. They came to her in the half-dark of dawn, when her mother's breathing was shallow and the Cord lay heavy against her collarbone.

\textit{But not for her. Never for her.}

Bryn's thumb brushed sweat from Aris`` temple. The skin was paper-thin over bone, mapped with old scars like healed cracks in dry earth. ''Your time is soon, mother,`` Bryn said, voice flat as a stone. ''Don``t worry. I''ll do my duty. I``ll take your last breath. Your debt to the Cord will be paid.''

Aris`` eyes flickered---just a shadow crossing water---and she turned her head. Not fear. Sadness. Deep and quiet, like a well long drained. ''Help me up,`` she whispered. ''I want to see the sunset over the Black River.``

Bryn shifted her weight, ready to lift. They moved slow, Aris leaning hard on Bryn's shoulder, each step a small surrender. The Cord's path wound through knee-high grass, silent except for the rasp of Aris'' breath and the distant cry of a lone crow. Bryn kept her hands loose at her sides, ready to steady or support. No words came. The Cord's weight against her collarbone was a familiar pressure---steady, expected.

At the river's edge, Aris stopped. She didn``t look at Bryn. Her gaze fixed on the water, where the sun bled gold into the valley's green folds. ''I tried,`` Aris said, voice thin as old thread. ''Several times. To escape. Every time\dots they brought me back.`` A pause. The wind stirred the reeds. ''You told them each time.``

Bryn's throat tightened. She hadn''t meant for Aris to know. Not like this. ``The Cord requires honesty,'' she said. ``Secrets weaken the weave.''

Aris didn``t nod. Just sighed---a sound like fabric tearing---and fell silent. They stood shoulder to shoulder, watching the light thin to amber, then rose. The river whispered over stones, indifferent.

Aris finally spoke, soft as a sigh. ''I don``t want to go to the Cord. The House Without Doors. Child\dots I have no friends there.'' Her hand lifted, trembling, to wipe her cheek. Came away wet. ``Everything I ever wanted\dots everything\dots is right here.''

Bryn's jaw locked. She``d said those words to dying women a hundred times. Duty was the only warmth left. ''You know the rules, Mama. And no tears. They``re\dots inefficient.'' Her voice sounded thin even to her own ears. She kept her hands unclenched at her sides, nails digging into palms.

Aris turned her face fully toward her. The sunset caught the tears tracking through the dust on her cheeks. ``You were always a good daughter,'' she said. Her thumb brushed Bryn's knuckle---barely there, gone before Bryn could flinch. ``And\dots more besides. I know the rules. And I have earned my tears.''

Bryn swallowed. The Cord felt tighter, a live thing against her throat. ``A breath, or a death, mother. The Cord demands. Nature chooses.''

Aris`` eyes were suddenly sharp, ancient in their weariness. ''No, Bryn.`` Her voice dropped to a blade's edge. ''No tears. \textit{And I choose}.`` Then she wasn''t leaning. She was falling.

Bryn's arms jerked out---too late, too weak---to catch air. Aris vanished over the edge. The river swallowed the sound of her fall before it could finish.

Bryn crashed to her knees. Stone bit through her robes. She pressed her palms flat to the earth, head down, and tried to break. Tried to let the grief crack her open like rotten wood. Nothing came. No sound. No wet heat on her cheeks. Just the hollow press of her own ribs, and the Cord---tight, insistent---pulsing once, twice, like a second heartbeat where hers should have been.

She stayed there, knuckles white in the dirt, waiting for tears that would not come. The Cord held her silent. The river kept singing. The sun dipped below the hills, and the light went gray.
\end{visionbox}

\vspace{1em}

The vision ends. The water ripples once, then is still. The gray thread has tied its ninth knot---a knot you did not make, in a pattern you do not recognize. You cut it with the edge of your silver mirror. The thread coils like a dead snake. You will burn it before dawn.

The locket is cold. The water is brackish. The mirror shows your face---weary, walnut skin cracked with age, the gray of your thinning hair indistinguishable from the gray of the thread. You do not recognize yourself. That is normal. The dead leave fingerprints on the living.

You set down the Symbol of Varnek Karn. You do not know what debt you have witnessed, only that it is not yours.

You write the names you heard: \emph{Bryn. Aris.} They mean nothing to you. Perhaps they will mean something to a future reader.

You are now a witness. That is enough.

\begin{flushright}
--- Saikou Ira
\end{flushright}

\vspace{1em}

\saikoumargin{I have performed this rite three times. The first time, I saw a mother fall into the Black River. The second time, I saw a daughter kneel and wait for tears that would not come. The third time, I saw a man in a ditch, his chest carved with the words that had condemned him. The locket is in a lead box in my study. The thread is still knotted. I have not untied the ninth knot. I am not ready for what it would release. --- S.I.}

