% chapter/chapter7.tex
% Adversary Templates & Quick Generators – Tools for the Weary Witness
% Rewritten in the voice of Lerris Fenwood, with marginalia from the Grey Wanderer,
% Saikou Ira, and Dusana of the Raven Road.

\epigraph{“The bestiary is a ledger. The templates are the columns. The generator is the hand that writes. Use them when the hour is late and the players have taken a road you did not expect.”}{—Lerris Fenwood, marginal note}

I have spent more hours than I care to count building adversaries from scratch. The work is good for the soul, but the soul is not always what the table needs. Sometimes it needs a bandit captain in thirty seconds. Sometimes it needs a boss whose phases fit on an index card. This chapter is for those times.

What follows are templates, generators, and procedures. They are not polished. They are **field tools**. Use them, bend them, discard them. The Grey Wanderer would approve. Saikou Ira would ask for a silver mirror. Dusana would ask for a story. I ask only that you count to eight before you roll.

\section*{Threat Level Scaling Table}

\noindent Every adversary in this compendium can be scaled by Threat Level (TL). Add the following adjustments to base stats. Use TL 0 for non‑threatening flavour, TL 1 for standard minions, TL 2 for elites, TL 3 for mini‑bosses, TL 4+ for major bosses (add one archetype trait from Chapter 1).

\begin{tabular}{|l|c|c|c|c|c|}
\hline
\textbf{Threat Level} & \textbf{Body} & \textbf{Wits} & \textbf{Spirit} & \textbf{Presence} & \textbf{Harm} \\
\hline
0 (flavour) & –1 & –1 & –1 & –1 & –2 \\
1 (base) & 0 & 0 & 0 & 0 & 0 \\
2 (elite) & +1 & +1 & +1 & +1 & +2 \\
3 (mini‑boss) & +2 & +2 & +2 & +2 & +4 \\
4+ (major boss) & +3 & +3 & +3 & +3 & +6 \\
\hline
\end{tabular}

\noindent\textbf{Example}: A TL 3 Bandit Captain (base Body 3, Harm 4) becomes Body 5, Harm 8, with +2 to all other attributes. Add one archetype trait (e.g., \emph{Cowardly} – flees if outnumbered).

\grey{Threat Level is a lie. The real threat is the story. But the lie helps the dice.}

\section*{Boss Design Checklist}

\noindent Use this checklist when building a custom boss. Keep the answer to each question to one sentence. If you cannot, the boss is too complicated.

\begin{enumerate}
    \item \textbf{Name \& Nature}: What is the boss, and what system (organ) of the story does it represent?
    \item \textbf{Phase Timer }: 6, 8, or 10? Write the phase names (e.g., \emph{Whisper} → \emph{Offer} → \emph{Claim}).
    \item \textbf{Three Moves}: One space‑control, one resource‑tax, one identity/knowledge twist.
    \item \textbf{Two Lair Actions}: Environmental toggles keyed to this boss’s lair.
    \item \textbf{Counters}: A courtesy, a tool, a sacrifice. Players should be able to discover these through play.
    \item \textbf{Drop}: A Diamond (see Chapter 1–6) or a bespoke boon tied to the organ.
\end{enumerate}

\noindent\textbf{Example (Abbreviated – The Ember Prefect from Chapter 6):}

\begin{verbatim}
1. Ember Prefect – schismatic inquisitor, represents the Church's hunger for purity.
2. Phase Timer [8]: Citation → Censure → Auto‑da‑fé → Purge.
3. Moves: Name reading (social), flame test (resource), purge (identity).
4. Lair: Candle Count (space), Shadow timer (resource).
5. Counters: relic provenance, crowd favour, rival sermon.
6. Drop: black ledger page, candle that never goes out.
\end{verbatim}

\saikou{I have used this checklist seven times. The seventh boss killed two PCs. I do not blame the checklist. I blame the dice. The dice do not apologise.}

\section*{Random Adversary Generator}

\noindent For when the players take a road you did not map. Roll 1d6 on each table. The result is not a full stat block – it is a **spark**. You provide the Threat Level, the Harm, and the fiction.

\subsection*{d6 – Creature Type}
\begin{enumerate}[label=\arabic*.]
    \item Mortal (humanoid, mercenary, cultist)
    \item Unquiet Dead (ghost, ghoul, poltergeist)
    \item Cursed (lycanthrope, vampire, slasher)
    \item Beast (natural or monstrous)
    \item Fae or Threshold Spirit
    \item Dungeon or Living Infrastructure
\end{enumerate}

\subsection*{d6 – Descriptor}
\begin{enumerate}[label=\arabic*.]
    \item Patient (waits, observes, follows)
    \item Hungry (attacks without warning, prefers the weak)
    \item Clever (sets traps, negotiates before fighting)
    \item Ancient (has a memory the party may need)
    \item Bound (cannot leave a location or object)
    \item Echo (repeats the last words of its victims)
\end{enumerate}

\subsection*{d6 – Motivation}
\begin{enumerate}[label=\arabic*.]
    \item Protects a threshold (door, ford, grave)
    \item Hungers for memory (names, faces, secrets)
    \item Seeks a witness (wants a confession or a story)
    \item Obeys a patron (unseen master, forgotten oath)
    \item Collects debts (silver, blood, promises)
    \item Simply wants to be left alone (will attack if provoked)
\end{enumerate}

\subsection*{d6 – Weakness}
\begin{enumerate}[label=\arabic*.]
    \item Silver (weapon or touch)
    \item Fire (or sunlight)
    \item A correct count (eight steps, eight breaths)
    \item A true name (spoken aloud with witness)
    \item A memory (offered freely, not taken)
    \item Fresh water (poured or splashed)
\end{enumerate}

\subsection*{d6 – Timer Size (for the encounter)}
\begin{enumerate}[label=\arabic*.]
    \item 4 segments (short, minor threat)
    \item 6 segments (standard)
    \item 8 segments (extended)
    \item 10 segments (epic, boss)
    \item Special: Timer is visible; players can tick it by fulfilling conditions
    \item Special: Timer ticks on player failures, not on time
\end{enumerate}

\noindent\textbf{Worked Example}: Roll 3, 2, 5, 4, 1 → Type: Cursed; Descriptor: Hungry; Motivation: Collects debts; Weakness: A true name; Timer : 4 segments. Result: A starving ghoul that once was a toll‑keeper. It demands payment from anyone who crosses its bridge. Its true name is carved on the underside of the keystone. The players have four segments to find the name or feed it a memory.

\grey{I rolled this generator once and got “Fae, Patient, Seeks a witness, A correct count, 6 segments.” The result was a Trow who asked every player to count to eight. One player sneezed on seven. The Trow took his shadow. He got it back after three sessions. The shadow was taller.}

\section*{Deploying Adversaries with Timers}

\noindent A timer is not a countdown. It is a **promise**. Use timers to make adversaries feel inevitable without being unfair. Below are common timer patterns.

\subsection*{Arrival Timer (Reinforcements)}
\noindent\textbf{Name}: \emph{Patrol Arrives [6]}. Ticks when the party makes noise, leaves evidence, or fails a Stealth roll. When full, 1d4 guards or a lieutenant arrives.

\subsection*{Doom Timer (Environmental)}
\noindent\textbf{Name}: \emph{Collapse [8]}. Ticks when the party fails a Body + Athletics roll or when the boss uses a specific lair action. When full, the ceiling falls (Harm 2 to everyone, terrain becomes Difficult).

\subsection*{Hunger Timer (Boss Enrage)}
\noindent\textbf{Name}: \emph{Bloodlust [6]}. Ticks each time the boss takes Harm. When full, the boss gains +1 die to all attacks and −1 die to defense (frenzy).

\subsection*{Debt Timer (Negotiation)}
\noindent\textbf{Name}: \emph{Debt to the Hollow [4]}. Ticks each time a character accepts a gift, breaks a promise, or lies. When full, a Oath‑Keeper appears and demands payment. The party can clear ticks by telling a true story or offering a memory.

\subsection*{Witness Timer (Social Pressure)}
\noindent\textbf{Name}: \emph{Crowd Turns [6]}. Ticks each time a character fails a social roll or is witnessed committing a crime. When full, bystanders become hostile or flee, and guards are summoned.

\saikou{A timer is a leash. The players hold the other end. If they forget, the leash tightens. I have seen a party ignore a \emph{Collapse} timer. The roof fell. They survived. The treasure did not.}

\section*{Quick Adversary Cards (Reference)}

\noindent The following cards are reproduced from Chapter 1 for quick reference. Print them, cut them, keep them behind the screen.

\vspace{1cm}

\noindent
\fbox{\parbox{0.45\textwidth}{\centering \textbf{Bandit Captain} (TL 2)\\ Body 3, Melee 2, Harm 4\\ \textit{Rabble Rouser} – On a Hit, rally 1d4 bandits.\\ \textit{Cowardly} – Flees if outnumbered.}}
\hfill
\fbox{\parbox{0.45\textwidth}{\centering \textbf{Ghostly Anchor} (TL 3)\\ Spirit 4, Lore 2, Harm 3\\ \textit{Unfinished Business} – Cannot be harmed until its anchor is addressed.\\ \textit{Bargain} – May offer a deal.}}

\vspace{1cm}

\noindent
\fbox{\parbox{0.45\textwidth}{\centering \textbf{Slasher} (TL 3)\\ Body 4, Melee 3, Harm 5\\ \textit{Regeneration} – Ignores first 2 Harm per scene.\\ \textit{Hunger Timer [6]} – When full, frenzy.}}
\hfill
\fbox{\parbox{0.45\textwidth}{\centering \textbf{Oath‑Keeper} (TL –)\\ Cannot be fought.\\ \textit{Demand} – One story, memory, or confession.\\ \textit{Weakness} – Pay the original debt.}}

\vspace{1cm}

\noindent
\fbox{\parbox{0.45\textwidth}{\centering \textbf{Thorn Courtier} (TL 3)\\ Presence 4, Sway 3, Harm 3\\ \textit{Courteous Snare} – A gift accepted is a debt owed.\\ \textit{Iron Offense} – Brandishing iron gives the GM 2 SB.}}
\hfill
\fbox{\parbox{0.45\textwidth}{\centering \textbf{Salt Prince Enforcer} (TL 2)\\ Body 3, Melee 2, Harm 4\\ \textit{Debt Marker} – Knows where you sleep.\\ \textit{Brine Curse} – Test Resolve or gain Salt‑Tongue.}}

\section*{Appendix for Chapter 7 – Quick Threat Level Reference}

\noindent Use this table to assign Harm, Fatigue, and key stats on the fly.

\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Threat Level} & \textbf{Typical Use} & \textbf{Base Harm} & \textbf{Base Fatigue} & \textbf{Key Attribute Range} \\
\hline
0 & Non‑combat, flavour & 1–2 & 1 & 1 \\
1 & Standard minion & 3 & 2 & 2–3 \\
2 & Elite, lieutenant & 4–5 & 3 & 3–4 \\
3 & Mini‑boss & 6–7 & 4 & 4–5 \\
4+ & Major boss & 8+ & 5+ & 5+ \\
\hline
\end{tabular}

\noindent\textbf{Example}: A TL 2 assassin (Body 3, Harm 4) has Body 3, Harm 4. Add +1 to all attributes for TL 2 → Body 4, Harm 4? Wait, the table says +1 to attributes and +2 to Harm. So Harm becomes 6. Consistent.

\lerris{I keep this table taped inside my GM screen. The tape is old. The table is worn. The numbers are still legible. That is enough.}

\section*{Final Advice from the Marginal Voices}

\noindent The Grey Wanderer: \emph{“Templates are not a substitute for attention. A bandit captain generated in thirty seconds will still kill a careless party. The dice do not care about your prep.”}

\noindent Saikou Ira: \emph{“I have faced a Oath‑Keeper generated from these tables. The generator said ‘Clock: 4 segments.’ The walker stayed for three sessions. The timer was a lie. The debt was real.”}

\noindent Dusana of the Raven Road: \emph{“The best adversary I ever ran was a hungry goat with a three‑timer. The players still tell stories about it. The generator did not make the goat. The story did.”}

\noindent Lerris Fenwood: \emph{“Use these tools. Then put them away. The adversary is not the template. The adversary is the witness. And the witness is always watching.”}

