\documentclass[11pt,letterpaper]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\usepackage{setspace}
\onehalfspacing
\usepackage{ebgaramond}
\usepackage[titles]{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{epigraph}
\setlength{\epigraphwidth}{0.75\textwidth}
\renewcommand{\epigraphflush}{flushright}
\renewcommand{\sourceflush}{flushright}

\usepackage{chapterfolder}
\newcommand{\cfile}[1]{\input{chapters/#1}}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[RE]{\textit{\leftmark}}
\fancyhead[LO]{\textit{\rightmark}}

\title{The Runic Warrior}
\subtitle{Codices of Flesh, Stone, and Steel}
\author{An Expansion for Fate's Edge}
\date{}

\begin{document}

\maketitle
\thispagestyle{empty}

\newpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
\textit{He knelt at the bank of the Black River. The water was cold. It smelled of iron and old rain. He opened the locket.}

\vspace{1em}

\textit{Inside was not a portrait. It was a braid of auburn hair, tied with a scrap of red thread.}

\vspace{1em}

\textit{This is not a story about a warrior who won a battle. It is a story about a warrior who learned to let go.}

\vspace{1em}

\textit{The thread held. The water carried. That is not a Rite. That is grace.}
\end{center}
\vspace*{\fill}
\newpage

\tableofcontents
\newpage

\part{Introduction and Foundations}

\chapter{The Warrior's Codex}
\cfile{introduction/01_what_is_body_runekeeper}
\cfile{introduction/02_philosophy_of_physical_obligation}
\cfile{introduction/03_how_to_use_this_expansion}

\chapter{The Runic Warrior's Toolkit}
\cfile{toolkit/01_codex_as_flesh_stone_steel}
\cfile{toolkit/02_thiasos_as_ally}
\cfile{toolkit/03_obligation_as_physical_weight}
\cfile{toolkit/04_new_core_talents}

\part{The Patrons of the Body}

\chapter{The Vowed (Oaths, Law, and Judgment)}
\cfile{patrons/vowed/01_oath_of_flame_and_light}
\cfile{patrons/vowed/02_oath_of_mercy_and_grace}
\cfile{patrons/vowed/03_mykkiel}
\cfile{patrons/vowed/04_inquisitor_prime}
\cfile{patrons/vowed/05_the_witness}

\chapter{The Threshold (Boundaries, Wards, and Passage)}
\cfile{patrons/threshold/01_the_sealed_gate}
\cfile{patrons/threshold/02_the_traveler}
\cfile{patrons/threshold/03_the_pale_shepherd}

\chapter{The Wild (Nature, Seasons, and the Hunt)}
\cfile{patrons/wild/01_grimmir}
\cfile{patrons/wild/02_the_carrion_king}
\cfile{patrons/wild/03_the_savage_heart}

\chapter{The Deep (Pressure, Storms, and the Abyss)}
\cfile{patrons/deep/01_khemesh}
\cfile{patrons/deep/02_xhakthul}
\cfile{patrons/deep/03_raeyn}

\chapter{The Shadow (Secrecy, Silence, and the Unseen)}
\cfile{patrons/shadow/01_ikasha}
\cfile{patrons/shadow/02_the_silent_choir}
\cfile{patrons/shadow/03_aveh}

\chapter{The Web (Fate, Binding, and Transformation)}
\cfile{patrons/web/01_inaea}
\cfile{patrons/web/02_isoka}
\cfile{patrons/web/03_malachai}

\chapter{The Forge (Creation, Iteration, and Refinement)}
\cfile{patrons/forge/01_clockwork_monad}

\chapter{The Grave (Ancestors, Debt, and Memory)}
\cfile{patrons/grave/01_varnek_karn}
\cfile{patrons/grave/02_the_high_mother}

\chapter{The Fracture (Knowledge, Paradox, and the Unknowable)}
\cfile{patrons/fracture/01_the_ninth}
\cfile{patrons/fracture/02_lunara}

\chapter{The Lethai Ghe'hai (Elite Operatives)}
\cfile{patrons/lethai/01_ghehai_syncretic}

\part{Mechanics and Options}

\chapter{New Talents for Runic Warriors}
\cfile{mechanics/01_foundation_talents}
\cfile{mechanics/02_advanced_talents}
\cfile{mechanics/03_prestige_talents}
\cfile{mechanics/04_patron_specific_talents}

\chapter{New Rites by Patron}
\cfile{mechanics/01_rites_vowed}
\cfile{mechanics/02_rites_threshold}
\cfile{mechanics/03_rites_wild}
\cfile{mechanics/04_rites_deep}
\cfile{mechanics/05_rites_shadow}
\cfile{mechanics/06_rites_web}
\cfile{mechanics/07_rites_forge}
\cfile{mechanics/08_rites_grave}
\cfile{mechanics/09_rites_fracture}
\cfile{mechanics/10_rites_lethai}

\chapter{The Thiasos as Ally}
\cfile{thiasos/01_core_rules}
\cfile{thiasos/02_living_thiasos}
\cfile{thiasos/03_sentient_thiasos_optional}
\cfile{thiasos/04_sentient_thiasos_by_patron}
\cfile{thiasos/05_sentient_thiasos_talents}

\chapter{The Codex as Object}
\cfile{codex/01_core_rules}
\cfile{codex/02_codex_by_patron}
\cfile{codex/03_codex_and_thiasos_unified}

\chapter{Runic Warrior Archetypes in Play}
\cfile{archetypes/01_dawn_warrior}
\cfile{archetypes/02_bloodsworn}
\cfile{archetypes/03_ghehai_operative}
\cfile{archetypes/04_beast_warrior}
\cfile{archetypes/05_fractal_warrior}
\cfile{archetypes/06_hybrid_builds}

\part{GM Guidance}

\chapter{Running Runic Warriors}
\cfile{gm/01_npcs}
\cfile{gm/02_faction_politics}
\cfile{gm/03_designing_custom_thiasos}
\cfile{gm/04_obligation_as_physical_scar}
\cfile{gm/05_sample_npcs_by_family}

\part{Appendices}

\appendix

\chapter{The Black River at Dawn}
\cfile{appendices/01_black_river_at_dawn}

\chapter{Quick Reference Tables}
\cfile{appendices/02_table_patrons_by_archetype}
\cfile{appendices/03_table_patrons_by_culture}
\cfile{appendices/04_table_talents_by_tier}
\cfile{appendices/05_table_rites_by_patron_and_tier}
\cfile{appendices/06_table_thiasos_by_patron}
\cfile{appendices/07_table_codex_by_patron}
\cfile{appendices/08_table_sentient_thiasos_by_patron}

\chapter{Glossary of Terms}
\cfile{appendices/09_glossary}

\chapter{Faction Relationship Matrix}
\cfile{appendices/10_faction_matrix}

\newpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
\textit{The thread held. The water carried. That is not a Rite. That is grace.}

\vspace{2em}

\textbf{The Runic Warrior: Codices of Flesh, Stone, and Steel}

\vspace{1em}
An Expansion for Fate's Edge
\end{center}
\vspace*{\fill}

\end{document}
