\section{Acasia --- ``Broken Marches''}
\label{chap:acasia}

\begin{tcolorbox}[colback=black!5,colframe=blue!60!black,title={Region Diagnostic Summary}]
\paragraph*{Current Hook}
The Empire conquered it first. The Empire left it last. Now neither remembers why.

\paragraph*{Core Conflict}
Players are caught between the \textbf{Margravine's} iron law, the \textbf{Lame King's} desperate legitimacy, and the \textbf{Curse} that devours broken promises.

\paragraph*{The One Thing Only This Region Does}
\textbf{Geometric Haunting:} Roads loop, bridges demand tolls in memory, and contracts signed on bridges become physically binding curses.

\paragraph*{The Ninth Taboo Manifestation}
In any Acasian scene, the ninth toll paid at a bridge is never the last. The wise pay eight and barter for the ninth. The ninth toll is collected by a child with no name. The child is the Curse's accountant.

\paragraph*{The Price}
Players sacrifice \textbf{memory, identity, or future labor} to gain power. Every reward requires a gamble: memory loss, Hollow attention, Obligation, or Corruption.

\paragraph*{Faction Triad}
\begin{tabular}{|p{0.25\linewidth}|p{0.25\linewidth}|p{0.25\linewidth}|p{0.15\linewidth}|}
\hline
\textbf{Faction} & \textbf{Goal} & \textbf{Method} & \textbf{Why players care} \\
\hline
\textbf{Margravine of the Broken March} & Keep the province from fracturing completely & Iron law, feud tallies, blood-oaths & She holds the only valid Imperial seals \\
\hline
\textbf{The Lame King} & Be recognized as a true sovereign & Traveling throne, debtor armies, decrees & He controls the roads the Margravine cannot \\
\hline
\textbf{Free Companies} & Find a permanent contract before winter & Selling loyalty, swapping colors mid-march & They are the only law between the claimants \\
\hline
\end{tabular}
\end{tcolorbox}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The Empire conquered it first. The Empire left it last. Now neither remembers why.}

\begin{quote}
  \textbf{Margravine of the Broken March (Elite):}  
  ``The tithe must be paid, not merely offered. A promise written in ash and sealed with iron will is worth more than a thousand gold coins scattered on the wind. The Empire is dead. Its weight is not.''
\end{quote}

\begin{quote}
  \textbf{Hedge-Witch (Commoner):}  
  ``Don't cross the Pale Causeway at dusk, especially not if you've got a silver tooth. The bridge'll remember it, and next time you pass, it'll want more than just your footfall. In Acasia, every road has a memory. The question is who pays the toll. The Empire forgot. That's why it fell.''
\end{quote}

\textbf{Genre / Mood:} Low-fantasy wasteland, mercenary realism, cursed borderlands.

\textbf{Starting Location:} The Pale Causeway's last inn, walls soot-blackened by coal and a century of gossip. A bounty poster flutters beside a half-empty hearth. The innkeeper is a one-eyed woman who pours small beer without asking. A stranger at the bar offers a sealed contract: retrieve a dowry chest from a Blackwood matriarch before the Margravine's feast day. ``The chest contains a claim, not coin,'' the stranger whispers. ``But a claim in Acasia is worth more than gold---it is a key to a bridge that does not loop. The Empire's surveyors marked it. Then they died.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Acasia was the first province the Utaran Empire swallowed when it turned its eyes east. The legions marched through the Pale Causeway, crushed the hill-kings, and planted their eagles in the Blackwood. For centuries, Acasia was the empire's eastern breadbasket and its first line of defense. Grain flowed west; coin flowed east. The roads were straight, the bridges paid, and the Curse was just a farmer's whisper.

Then the empire rotted from within. When the last Utaran emperor died and the provinces broke away, Acasia lingered for another hundred years --- a ghost province whose governors still stamped Utaran seals, whose courts still cited Utaran law, whose coins still bore the emperor's face. No one came from the west to collect the taxes. No one sent legions to enforce the peace. The Acasians kept the forms, because they had never learned to keep anything else. They had no native king, no unifying faith, no common tongue beyond the legionary's Latin. They were the empire's children, and when the empire died, they became orphans.

For a century they clung to the ruins. Then the plagues came. Then the famines. Then the Curse --- a slow rot of roads that led nowhere, bridges that looped back on themselves, villages that woke to find their neighbors' houses standing empty. The old imperial surveys became maps of places that no longer existed. The grain blights turned the storehouses to dust. The free companies, once hired to keep the peace, became the only law.

That was two hundred years ago.

Now Acasia is a wound that will not close. It sits between Ecktoria to the west and Vhasia to the east, claimed by both, ruled by neither. Petty kings raise their banners in burned keeps. Free Companies trade loyalty like grain, swapping colors mid-march. The Margravine of the Broken March keeps a tally of feuds, binding bloodlines to her feast days. The Lame King travels on a throne carried by debtors, making law by pointing. And the Curse --- the nameless thing that makes crossroads turn back on themselves --- is not a myth. It is the weight of broken promises, and Acasia is drowning in them.

South, on the marshy islands of the Amaranthine coast, the city of Silkstrand squats on reclaimed silt, trading in silk, silver, and secrets. The Acasians call it ``the Sink'' --- a place where the Curse thins, but where law is measured in coin, not custom. Few Acasians go there. Fewer return unchanged.

\vspace{2mm}
\textbf{What you notice first:} The smell of mildew on old parchment and wet ash. The way every milestone has been recarved --- first in Utaran, then in a half-dozen failed scripts. The silence of a village that heard you coming and chose not to open its gates.

\textbf{The one rule that kills newcomers:} Never sign a contract on a bridge. The river is a witness, and it never forgets a signature. The Empire's last governor signed a treaty on the Pont-du-Tithe. No one knows what it said. The bridge remembers.
\end{tcolorbox}

\subsection*{Regional Mood: Famine, Ambition, and the Ghost of Empire}

Here, the past is not a memory --- it is a geometry. The Utaran Empire left behind roads that lead to nowhere, laws that contradict each other, and a hunger that no harvest can fill. Acasia's petty kings hold court in burned keeps, their crowns forged from iron nails pulled from famine coffins. The Margravine keeps a tally of feuds, binding bloodlines to her feast days. Free Companies trade loyalty like grain, swapping colors mid-march. And beneath it all, the Curse waits. It is not a monster. It is a wrong turn: the bridge that leads back to the same shore, the milestone that faces the wrong direction, the bell that tolls for no funeral. In Acasia, the land itself has become a broken promise, and the fine print is written in mildew.

The Curse is not ancient. It is a legacy. Some say it began with the last Imperial Survey, when the surveyors tried to map every bridge, every ford, every toll---and in the measuring, broke something. Others say it was the plagues, when the dead were buried in unmarked pits and the priests forgot the proper rites. The hedge-witches say it was the empire's own fault: they ruled Acasia for centuries but never taught the Acasians to rule themselves. When the empire left, the land had no king, no law, no memory. The Curse filled the gap.

Now the Margravine and the Lame King fight over the scraps, while free companies sell their swords to whichever claimant pays first. The hedge-witches tie red thread around bridges and leave offerings of salt at crossroads, hoping to keep the Curse from spreading. The villagers pray to saints the Empire forgot. And the roads --- the old imperial roads --- still march west toward a capital that no longer exists. The capital does not answer. But the roads keep marching.

\subsection*{The Curse of Acasia (Campaign Clock)}
The Curse is a slow accumulation of broken promises and unpaid tolls. Track it with a visible timer:

\textbf{Curse Awakening [10]} – Advances when:
\begin{itemize}
\item A contract is broken on a bridge (+1)
\item A character speaks the ninth word in a negotiation (+1)
\item The party ignores a road omen (GM spends 1 SB to tick +1)
\item A free company flips allegiance mid-contract (+2)
\end{itemize}
At 4 segments: Navigation rolls in Acasia gain +1 DV (roads shift).  
At 8 segments: The Curse Stirrs complication triggers once per session (roll on the table below).  
At 10 segments: The Curse manifests as a physical threat --- a bridge eats a village, a road swallows a patrol, or the ghost of the last emperor rides.

\textbf{Curse Stirrs (d6)}  
When the timer reaches 8, roll 1d6 for an immediate complication:
\begin{enumerate}
\item A milestone points the wrong way. Navigation DV +2 for the scene.
\item A bridge you crossed earlier now loops back to where you started. Lose 1 segment of progress.
\item A toll collector's ghost demands payment in a currency that no longer exists. Pay 1 Boon or mark 1 Fatigue.
\item A child with no shadow offers directions. If you follow, you arrive at your destination --- but you forget why you came.
\item The Margravine's seal cracks. One of your Diamonds (writ, charter, token) becomes a forgery until you re-witness it.
\item The Hollow's breath freezes the ford. Crossing requires a Desperate roll; on a Miss, you lose 1 Supply (drowned).
\end{enumerate}

\subsection*{Prominent NPCs of Acasia}
\label{sec:acasia-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Zara \textit{``the Ledger''} & Margravine of the Broken March & Keep the province from fracturing completely & She has never lost a feud; her sword is also a counting rod. She claims descent from the last Utaran governor. \\
\hline
Oren \textit{``Tin Crown''} & The Lame King & Be recognized as a true sovereign & His traveling throne is carried by debtors; he pays them in promises. His grandfather was a free company captain who declared himself king. \\
\hline
Ysabeau & Hedge-Witch of the Blackwood Charcoalers & Survive the Curse and keep her kin from the rope & She knows which bridges eat travelers, and her price is a story that ends badly. Her grandmother witnessed the last imperial survey. \\
\hline
The Bride & Dowry chest claimant (Blackwood matriarch's heir) & Recover the chest before the Margravine's feast day & She carries a dagger in her wedding veil. The claim is to a bridge that drowned. The bridge was built by the Utaran legions. \\
\hline
Vladan & Free Company Captain (``The Rusted Banner'') & Find a permanent contract before winter & He has served three kings this year. He cannot remember their names. His company's banner is an Utaran eagle, torn and stitched. \\
\hline
Sister Ash & Ex-Imperial surveyor & Finish the old map of Acasia & She visits villages that no longer exist. The map shows them anyway. She carries a sealed chest of the last governor's papers. \\
\hline
\end{longtable}

\subsection*{Names of Acasia}
\label{sec:acasia-names}

Inspired by the martial duchies of a fallen southern empire: clipped Latin, war-names, and the hard consonants of mountain keeps. Many families still bear Utaranized surnames, though they have not spoken the language in generations.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Alboin & Theodelinda & da Ponte & \textit{the Stiff} \\
\hline
Authari & Rosamund & del Ferro & \textit{the Toll-Fist} \\
\hline
Liutprand & Gundiberga & di Rocca & \textit{Bridge-Born} \\
\hline
Desiderius & Anspranda & Longobardus & \textit{Blackwood} \\
\hline
Ratchis & Veleda & Saxo & \textit{the Unremembered} \\
\hline
Aistulf & Gisela & Capoferro & \textit{Ash-Finger} \\
\hline
\end{longtable}

\paragraph*{(Pass/Ruin/Toll-Town)} Broken milestone on the old Imperial Road; borders ``moved'' overnight. Vine-terrace hillside with an abandoned press; strangler ivy owns it now.

\section*{\spadecard{A} --- Places}
\label{sec:acasia-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Broken Milestone} \hfill \spadecard{2} \textit{[RUIN] [OMEN]} \\
  On the old Imperial Road; borders ``moved'' overnight. \\
  \textit{The stone has been chiseled seven times. The last mark is still wet. The first was cut by Utaran legionaries.} \\
  \textbf{Mechanical hook:} If you camp here, roll Spirit + Lore (DV 4). On success, you learn the true border line (+1 die on navigation for this leg). On failure, the milestone moves while you sleep --- start a \textbf{Lost Ways [4]} timer.

\item \textbf{Vine-Terrace Hillside} \hfill \spadecard{3} \textit{[RUIN] [HARVEST]} \\
  Abandoned press; strangler ivy owns it now. \\
  \textit{The ivy has grown around a skeleton. The skeleton still holds a wine cup. The vintage was the year the last governor left.} \\
  \textbf{Mechanical hook:} Foraging here gives 1 Supply, but the ivy is possessive. On a Miss, you are entangled (Body+Athletics DV 3 to break free) and tick \textbf{Curse Awakening +1}.

\item \textbf{Toll-Bridge Town} \hfill \spadecard{4} \textit{[TOLL] [BOUNDARY]} \\
  Over a cold river; two tolls, no receipts. \\
  \textit{The second toll is collected by a child. The child does not speak. The bridge was built by the legions. It remembers their footsteps.} \\
  \textbf{Mechanical hook:} Crossing costs 1 Boon or 1 Supply. If you refuse, the child writes your name in chalk. The bridge will demand double next time.

\item \textbf{Wolfstairs Pass} \hfill \spadecard{5} \textit{[MOUNTAIN] [AMBUSH]} \\
  Switchbacks beneath a stonefall cliff. \\
  \textit{The stones were placed by a landslide that knew your name. The pass was the empire's eastern gate. No army has crossed it in a century.} \\
  \textbf{Mechanical hook:} Navigation DV +2. On a Partial, you are spotted by a free company scout --- start \textbf{Rival Hunt [4]}.

\item \textbf{Sootfall Abbey Ruins} \hfill \spadecard{6} \textit{[RUIN] [DEATH]} \\
  Bell tower intact; crypts breathing. \\
  \textit{The breathing is a prayer. The monks are still down there. They have been praying for the empire's soul for two hundred years.} \\
  \textbf{Mechanical hook:} A character who prays with the monks (Spirit+Lore DV 3) clears 1 Fatigue, but the monks ask for a story about a broken promise. Refusal ticks \textbf{Curse Awakening +1}.

\item \textbf{Hill-Motte} \hfill \spadecard{7} \textit{[FORTIFICATION] [WAR]} \\
  Fresh palisade; ditch scattered with caltrops. \\
  \textit{The caltrops are from a siege that ended last week. The bodies are still rotting. The siege was over a bridge toll. The bridge has not collected in fifty years.} \\
  \textbf{Mechanical hook:} The lord of the motte will hire the party as mercenaries (1 Supply). If they refuse, he accuses them of being spies --- start \textbf{Feud Spark [4]}.

\item \textbf{Border-Stone Ring} \hfill \spadecard{8} \textit{[BOUNDARY] [OMEN]} \\
  Carved with seven crowns; each points wrong. \\
  \textit{The crowns belong to kings who never existed. The stones remember them anyway. The eighth crown was the empire's. It is not carved. It is implied.} \\
  \textbf{Mechanical hook:} A character who touches the eighth (empty) crown must test Resolve (DV 4). On failure, they gain the Condition \textit{Marked by the Curse} (GM may spend 1 SB to worsen Position in any negotiation).

\item \textbf{Blackwood Charcoalers' Hollow} \hfill \spadecard{9} \textit{[FOREST] [WITCHCRAFT]} \\
  Witch-posts at every path. \\
  \textit{The witch-posts were made from the gibbets of tax collectors. The tax collectors still swing. They were collecting for the empire. They have not stopped.} \\
  \textbf{Mechanical hook:} A hedge-witch offers to read your fortune (1 Boon). On a success, she warns of the next Curse Stirrs trigger. On a failure, she demands a memory (erase a minor fact).

\item \textbf{Salt-Road Ford} \hfill \spadecard{10} \textit{[RIVER] [DEATH]} \\
  Rains expose old bones in the chalk banks. \\
  \textit{The bones are from three wars. They are not all human. The ford was the empire's salt route. The salt is still in the water.} \\
  \textbf{Mechanical hook:} Crossing costs 1 Fatigue (cold). If you drink the water, test Endurance DV 3. On failure, you see the ghosts of the drowned for the next scene (–1 die on all actions).

\item \textbf{Iron Mine Adits} \hfill \spadecard{J} \textit{[UNDERGROUND] [WHISPERS]} \\
  Held by a miners' commune; the air is full of whispers. \\
  \textit{The whispers are the mine's memory. It remembers every collapse. The last collapse was the year the empire fell. The miners are still digging.} \\
  \textbf{Mechanical hook:} The commune will trade 1 Supply for a miner's story. If the story is true, they also reveal a hidden exit --- treat the next travel timer as 2 segments shorter.

\item \textbf{Margravine's Hunting Lodge} \hfill \spadecard{Q} \textit{[NOBLE] [SECRETS]} \\
  Tapestries of victories that never were. \\
  \textit{The tapestries were woven by a blind weaver. The threads are from her own hair. The victories are Utaran. The weaver was Acasian.} \\
  \textbf{Mechanical hook:} A character who studies the tapestries (Wits+Lore DV 4) learns a noble's secret (a String). On a failure, the blind weaver's ghost asks for a favor --- start \textbf{Weaver's Debt [4]}.

\item \textbf{War-Camp City} \hfill \spadecard{K} \textit{[WAR] [REFUGE]} \\
  Tents around a burned keep; every banner claims the throne. \\
  \textit{The keep was burned by its own lord. He did not want to share the heat. The lord was the last imperial governor. He died in the fire.} \\
  \textbf{Mechanical hook:} To enter the camp, you must swear a temporary oath to a claimant. Breaking it ticks \textbf{Curse Awakening +2}.

\item \textbf{The Pale Causeway} \hfill \spadecard{A} \textit{[ROAD] [CURSE]} \\
  The last high road that survives the spring thaws. \\
  \textit{The causeway is made of bones. The bones are from the workers who built it. The workers were Acasian. The empire paid them in promises.} \\
  \textbf{Mechanical hook:} Crossing the Causeway at dusk forces a Resolve test (DV 4). On failure, you gain the Condition \textit{Debtor to the Dead} (any spirit can demand one service from you). On a critical failure, the Curse Awakening timer ticks +2 immediately.
\end{enumerate}

\paragraph*{(Petty Crowns/Priests/Companies)} Tithe-collector's runner with a tally-rod and an empty stomach; roadside prior with three lay brothers guarding a relic.

\section*{\heartcard{A} --- People \& Factions}
\label{sec:acasia-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Tithe-Collector's Runner} \hfill \heartcard{2} \textit{[AUTHORITY] [HUNGER]} \\
  Tally-rod, empty stomach, and a list of names. \\
  \textit{His tally-rod has notches from a dozen villages. The villages are now empty. He collects for the Margravine. She does not ask where the grain goes.} \\
  \textbf{Debt they owe:} He owes his soul to the Margravine's ledger. If you give him a name from the list, he will mark it paid --- but the ghost of that name will haunt you. \\
  \textbf{Price of aid:} 1 Supply or a story about a tax collector who died hungry.

\item \textbf{Roadside Prior} \hfill \heartcard{3} \textit{[FAITH] [RELIC]} \\
  Three lay brothers guarding a relic. \\
  \textit{The relic is a finger bone. The prior claims it belonged to a saint who starved to death. The saint was an Utaran missionary. The Acasians ate his offerings.} \\
  \textbf{Debt they owe:} The prior's order swore to feed the poor. They have failed. He will help anyone who shares a meal. \\
  \textbf{Price of aid:} A crust of bread and a prayer. If you refuse, the relic's ghost will follow you (–1 die on all stealth rolls for one scene).

\item \textbf{Hedge-Witch} \hfill \heartcard{4} \textit{[WITCHCRAFT] [KNOWLEDGE]} \\
  Knows which bridges eat travelers. \\
  \textit{Her price is a story. The story must have a bad ending. She has heard all the good ones. The bridges listen when she speaks.} \\
  \textbf{Debt they owe:} She once cursed a tax collector. The curse still echoes. If you help her lift it, she will teach you a route that ignores the Curse once. \\
  \textbf{Price of aid:} A story that ends with a death. If the story is true, she gives you a Salt Ward (Diamond). If false, the bridges will remember you.

\item \textbf{Free Company Captain} \hfill \heartcard{5} \textit{[MERCENARY] [FATIGUE]} \\
  Pike and shot between contracts. \\
  \textit{He has served three kings this year. He does not remember their names. His first contract was with a man who called himself emperor. The man is dead. The contract is not.} \\
  \textbf{Debt they owe:} He owes back pay to his company. If you pay 1 Supply, he will fight for you for one scene. \\
  \textbf{Price of aid:} A contract signed in blood. If you break it, the Curse Awakening ticks +2.

\item \textbf{River Reeve} \hfill \heartcard{6} \textit{[LAW] [CORRUPTION]} \\
  Rents every boat twice. \\
  \textit{His paper is a single page folded into a boat. The other half is elsewhere. The other half is in the Margravine's counting house. She does not know.} \\
  \textbf{Debt they owe:} He stole the half-page from the Margravine. If you return it, he will waive all tolls for a season. \\
  \textbf{Price of aid:} A bribe of 1 Boon or a secret about the river. He will also betray you for a better offer.

\item \textbf{Salt-Baron} \hfill \heartcard{7} \textit{[TRADE] [VIOLENCE]} \\
  Hired blades; owns the ford and your timetable. \\
  \textit{His salt is gray. It was cut with ashes from a burned village. The village refused to pay the imperial salt tax. The empire is dead. The salt is not.} \\
  \textbf{Debt they owe:} He took a loan from the Margravine's counting house. If you pay it (2 Supply), he will give you a Salt Charter (pass any ford free). \\
  \textbf{Price of aid:} A wagon of gray salt. Using it in a bargain gives +1 die to intimidation, but the ghost of the burned village will haunt the transaction.

\item \textbf{Blackwood Matriarch} \hfill \heartcard{8} \textit{[NOBLE] [VENGEANCE]} \\
  Tends the feud like a garden. \\
  \textit{Her garden grows tombstones. She waters them with tears---her enemies'. Her grandmother was the last Acasian to hold an imperial title. The empire took it back. She has not forgotten.} \\
  \textbf{Debt they owe:} The empire stole her grandmother's title. If you recover the original charter (in the Sealed Dowry Chest), she will name you her heir. \\
  \textbf{Price of aid:} A lock of hair from a blood enemy. She will not say whose.

\item \textbf{Ex-Imperial Surveyor} \hfill \heartcard{9} \textit{[KNOWLEDGE] [LONELINESS]} \\
  Has the last accurate map. \\
  \textit{The map shows villages that no longer exist. The surveyor visits them anyway. She leaves offerings of bread and ink. The ghosts do not eat. They watch.} \\
  \textbf{Debt they owe:} She lost the map's key --- a brass compass. Find it, and she will draw you a route that ignores the Curse for one journey. \\
  \textbf{Price of aid:} A true name of a dead village. She will trade a secret for it.

\item \textbf{``King'' of Three Villages} \hfill \heartcard{10} \textit{[CLAIMANT] [IRON]} \\
  Iron-nail crown, iron-will taxes. \\
  \textit{His crown was a plowshare. He reversed the curse. The curse reversed back. He does not know.} \\
  \textbf{Debt they owe:} He borrowed the plowshare from a hedge-witch. The witch wants it back. Return it, and he will grant you a tax waiver. \\
  \textbf{Price of aid:} A day's labor in his fields. If you refuse, he accuses you of oath-breaking --- start a \textbf{Feud Spark [4]}.

\item \textbf{Bride with No Dowry but a Claim} \hfill \heartcard{J} \textit{[CLAIM] [DESPERATION]} \\
  Daggers in the wedding chest. \\
  \textit{The claim is to a bridge that drowned. The bride intends to collect. The bridge was built by the legions. The legions are gone. The bridge is not.} \\
  \textbf{Debt they owe:} Her dowry was the bridge's toll. The Margravine seized it. If you help her collect, she will give you a \textbf{Bridge Token} (Diamond). \\
  \textbf{Price of aid:} A dagger from her chest. Using it in a duel gives +1 die, but the bridge's ghost will claim the loser.

\item \textbf{Margravine of the Broken March} \hfill \heartcard{Q} \textit{[AUTHORITY] [FEUD]} \\
  Half saint, half tax roll. \\
  \textit{She has never lost a feud. Her tally stick is her sword. Her great-great-grandfather was the last imperial governor's clerk. He stole the seal. She still has it.} \\
  \textbf{Debt they owe:} She is losing her grip on the province. If you settle a feud for her, she will give you a \textbf{Blood-Peace Charter} (Diamond). \\
  \textbf{Price of aid:} A signature on her tally stick. Signing binds you to her service until the feud is settled.

\item \textbf{The Lame King} \hfill \heartcard{K} \textit{[CLAIMANT] [FATIGUE]} \\
  Traveling throne; makes law by pointing. \\
  \textit{His throne is a sedan chair carried by debtors. They owe him everything. He owes the Margravine a bridge toll. He has not paid. The bridge remembers.} \\
  \textbf{Debt they owe:} He cannot pay the toll. If you pay it for him (2 Supply), he will name you his heir for a day --- you may issue one royal decree. \\
  \textbf{Price of aid:} A ride on his throne. The debtors will carry you, but you will owe them a favor in return.

\item \textbf{The Cursed Child of Silkstrand} \hfill \heartcard{A} \textit{[MYTH] [OMEN]} \\
  A rumor more than a person; her laughter ends sieges. \\
  \textit{She is said to be twelve years old. She has been twelve for a century. She was born on the day the last imperial governor died. She does not age. She does not forget.} \\
  \textbf{Debt they owe:} She is the Curse's accountant. If you find her, she will tell you how many bridges you owe. \\
  \textbf{Price of aid:} A name you have never spoken. She will write it in her ledger. The Curse will remember it forever.
\end{enumerate}

\paragraph*{(Feud/Levy/Weather/Curse)} Peat-fog; all horns sound like someone calling your name. Sudden levy---every traveler pressed for a day's service. Bridge feud; rival banners drop the chain on your cart.

\section*{\clubcard{A} --- Complications/Threats}
\label{sec:acasia-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peat-Fog} \hfill \clubcard{2} \textit{[WEATHER] [FEAR]} \\
  All horns sound like someone calling your name. \\
  \textit{The name is not yours. Answer it, and the fog will take you. The fog is the empire's last breath.} \\
  \textbf{Mechanical hook:} Navigation rolls are Desperate for the scene. On a Miss, a ghostly tax collector appears. It demands a story or a memory. Refusal ticks \textbf{Curse Awakening +1}.

\item \textbf{Sudden Levy} \hfill \clubcard{3} \textit{[AUTHORITY] [TIME]} \\
  Every traveler is pressed for a day's service. \\
  \textit{The levy master has a list. Your name is on it---in fresh ink. The list was copied from an imperial tax roll. The roll is two hundred years old.} \\
  \textbf{Mechanical hook:} Pay 1 Supply or serve. Serving forces a Resolve test (DV 4). On failure, you gain the Condition \textit{Impressed} (–1 die to all actions until you escape). Start a \textbf{Levy Escape [4]} timer.

\item \textbf{Bridge Feud} \hfill \clubcard{4} \textit{[BOUNDARY] [VIOLENCE]} \\
  Rival banners drop the chain on your cart. \\
  \textit{The chain was forged from a broken oath. It rattles when liars approach. The oath was sworn to the emperor. The emperor is dead. The chain does not care.} \\
  \textbf{Mechanical hook:} The chain will only lift if you swear an oath to one of the feuding parties. Swearing ticks \textbf{Curse Awakening +1} (oaths made on bridges always compound). Refusing forces a fight --- treat as a Cap 2 adversary.

\item \textbf{Grain Blight} \hfill \clubcard{5} \textit{[HARVEST] [FEAR]} \\
  Doubles the tithe; refusal means torches at night. \\
  \textit{The blight has a face. It is the face of a tax collector who died hungry. The collector worked for the empire. The empire did not pay him.} \\
  \textbf{Mechanical hook:} You may pay double (2 Supply) or refuse. Refusing starts a \textbf{Village Hunt [4]} timer. When full, the villagers attack at night (Cap 3 mob).

\item \textbf{Scree Slide} \hfill \clubcard{6} \textit{[MOUNTAIN] [BLOCKED]} \\
  Seals the pass; the ``safe'' detour is owned by your enemy. \\
  \textit{The enemy raised the toll yesterday. He was expecting you. The pass was the empire's eastern gate. It still remembers the legions.} \\
  \textbf{Mechanical hook:} Pay 2 Supply to the enemy or find another way. Searching for another way forces a Wits+Survival roll (DV 5). On a failure, you lose 1 segment of progress and gain a \textbf{Black Mark} with the enemy.

\item \textbf{Wedding Turns Ambush} \hfill \clubcard{7} \textit{[SOCIAL] [TRAP]} \\
  You are caught between two bloodlines. \\
  \textit{The bride's veil is a shroud. The groom's ring is a knife. The feud began when the empire divided the valley. The valley has not healed.} \\
  \textbf{Mechanical hook:} You must choose a side or flee. Choosing a side gives you an ally (Cap 2) but starts a \textbf{Feud Spiral [6]} timer with the other. Fleeing costs 1 Fatigue and ticks \textbf{Curse Awakening +1}.

\item \textbf{Witch's Tithe Night} \hfill \clubcard{8} \textit{[WITCHCRAFT] [DEBT]} \\
  Lights move in the woods; owe nothing or else. \\
  \textit{The lights have voices. They are arguing over who should collect. They are the ghosts of imperial tax collectors. They have not stopped working.} \\
  \textbf{Mechanical hook:} The ghosts demand a story. If you cannot tell one, they take 1 of your Boons (or 1 Supply). If you have nothing, they take a memory — lose 1 XP (refundable by a quest).

\item \textbf{Pox Sign} \hfill \clubcard{9} \textit{[DISEASE] [ISOLATION]} \\
  On a village gate; your escorts desert you. \\
  \textit{The pox is a rumor. The desertion is real. The pox came with the last imperial supply train. The supply train never arrived.} \\
  \textbf{Mechanical hook:} You must travel alone for the next leg (no assists). Start a \textbf{Disease Clock [4]}. When full, each PC gains 1 Fatigue.

\item \textbf{Condotta Breaks} \hfill \clubcard{10} \textit{[MERCENARY] [BETRAYAL]} \\
  The Free Company flips colors mid-march. \\
  \textit{The new colors are the same as the old colors. The captain just wanted to raise morale. He was hired by the Margravine. He was also hired by the Lame King. He is collecting both salaries.} \\
  \textbf{Mechanical hook:} The company now fights for your enemy. You may bribe them back (2 Supply) or fight them (Cap 3). Bribing ticks \textbf{Curse Awakening +1} (broken loyalty compounds).

\item \textbf{Heretic Preacher} \hfill \clubcard{J} \textit{[FAITH] [RIOT]} \\
  Sparks a march; toll-gates come down in splinters. \\
  \textit{The preacher's text is a tax receipt. He reads it like scripture. The receipt was issued by the empire. It is still valid. No one knows to whom.} \\
  \textbf{Mechanical hook:} The preacher will grant you safe passage through the riot if you give him a relic (any Diamond). If you refuse, the mob attacks (Cap 2, and you are separated from allies).

\item \textbf{Imperial Pretender} \hfill \clubcard{Q} \textit{[CLAIMANT] [POLITICS]} \\
  Arrives; every petty lord changes tabards. \\
  \textit{The pretender's seal is a sunburst. The sunburst is slightly crooked. The seal was stolen from the last governor's tomb. The governor's ghost is following him.} \\
  \textbf{Mechanical hook:} The pretender offers you a charter (Diamond) if you swear loyalty. Swearing ticks \textbf{Curse Awakening +2} (the ghost of the governor will remember). Refusing makes him a permanent enemy (start a \textbf{Pretender's Hunt [6]} timer).

\item \textbf{River Overruns the Levee} \hfill \clubcard{K} \textit{[WEATHER] [DESTRUCTION]} \\
  The pontoon goes; your rivals don't. \\
  \textit{The river's name is ``Debt.'' It remembers every unpaid crossing. The empire's engineers built the levees. They calculated the flood levels. They did not calculate the Curse.} \\
  \textbf{Mechanical hook:} You lose 1 Supply (washed away). If you have no Supply, you lose 1 Boon instead. The GM may also tick \textbf{Curse Awakening +1} (the river demands interest).

\item \textbf{The Curse Stirrs} \hfill \clubcard{A} \textit{[CURSE] [TRAP]} \\
  No matter the road, you return to the same crossroads. \\
  \textit{The crossroads has a signpost. The arms are labeled ``Regret,'' ``Hunger,'' ``The Same.'' The post was planted by the last imperial surveyor. She is still walking.} \\
  \textbf{Mechanical hook:} Roll on the \textbf{Curse Stirrs} table (see Regional Mood section). The party cannot leave until they pay a toll — a story, a memory, or a name.
\end{enumerate}

\paragraph*{(Papers/Claims/Safe-Conduct)} Toll-exemption plaque for one bridge (once); monastery letter for bed and bread on a named road.

\section*{\diamondcard{A} --- Rewards/Leverage}
\label{sec:acasia-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Toll-Exemption Plaque} \hfill \diamondcard{2} \textit{[TRAVEL] [ONCE]} \\
  For one bridge (once). \\
  \textit{The plaque is a bronze leaf. Hold it up, and the toll-keeper will spit---but step aside. The leaf was minted by the empire. The empire's mines are closed. The metal is still good.} \\
  \textbf{Curse cost:} Using the plaque ticks \textbf{Curse Awakening +1} (the bridge will remember your face). If you use it on a bridge that has already taken a toll, the plaque crumbles and the bridge's ghost follows you for a day.

\item \textbf{Monastery Letter} \hfill \diamondcard{3} \textit{[HOSPITALITY] [ONCE]} \\
  For bed and bread on a named road. \\
  \textit{The letter is sealed with a wax tear. The monks will honor it---or they will not. The abbot was an imperial clerk. He still follows the old forms.} \\
  \textbf{Curse cost:} The monks will ask you to pray for the dead. If you refuse, the letter burns and you gain a \textbf{Black Mark} with the monastery. If you pray, clear 1 Fatigue but tick \textbf{Curse Awakening +1} (the dead listen).

\item \textbf{Wine-Right} \hfill \diamondcard{4} \textit{[TRADE] [ONGOING]} \\
  On an abandoned terrace; locals will work for shares. \\
  \textit{The wine is from a vine that grows over a mass grave. It tastes of iron. The grave holds the last imperial garrison. They were paid in promises.} \\
  \textbf{Curse cost:} Drinking the wine gives +1 die to your next social roll, but the garrison's ghosts will appear in your dreams. The GM gains 1 SB to use on a nightmare complication.

\item \textbf{Condotta Contract} \hfill \diamondcard{5} \textit{[MERCENARY] [ONCE]} \\
  A signed pike contract (one battle on your terms). \\
  \textit{The contract is written in blood. The blood is from the captain's own finger. The contract was copied from an imperial legionary oath. The words have changed. The weight has not.} \\
  \textbf{Curse cost:} If the contract is broken, tick \textbf{Curse Awakening +2}. The captain's ghost will demand a new contract — signed in your own blood.

\item \textbf{Tithe-Remission Writ} \hfill \diamondcard{6} \textit{[AUTHORITY] [ONCE]} \\
  For a village; they owe you a season's labor. \\
  \textit{The writ is stamped with a bell. The bell's clapper is a finger bone. The bell was cast from melted imperial coinage. The coinage was debased. The bell rings true.} \\
  \textbf{Curse cost:} Using the writ forces the village to work for you, but the bell's ghost will follow you. Every season, you must offer it a story or lose 1 Boon.

\item \textbf{Border-Stone Adjustment} \hfill \diamondcard{7} \textit{[BOUNDARY] [ONCE]} \\
  Move a line two fields over. \\
  \textit{The adjustment is a lie. The truth is that the line was always there. The line was surveyed by the empire's best geometers. They are all dead. Their measurements are not.} \\
  \textbf{Curse cost:} The ghost of the surveyor will appear and demand you measure the new line with your own feet. Walk it exactly (8 steps) or lose 1 Fatigue and tick \textbf{Curse Awakening +1}.

\item \textbf{Pass-Key Charm} \hfill \diamondcard{8} \textit{[TRAVEL] [ONGOING]} \\
  Recognized by Pale Causeway watchmen. \\
  \textit{The charm is a knotted cord. Untie it to pass---or to stay. The cord was tied by an imperial courier. The courier's route was swallowed by the Curse. The cord remembers the way.} \\
  \textbf{Curse cost:} The charm works for one season. After that, the courier's ghost will visit you. He will ask for a message to deliver. If you refuse, the charm becomes a curse — all roads lead back to the same crossing.

\item \textbf{Sealed Dowry Chest} \hfill \diamondcard{9} \textit{[CLAIM] [MYSTERY]} \\
  Of claims, not coin; certain doors open. \\
  \textit{The chest is locked with a riddle. The answer is a name you do not know. The name is the last imperial governor's. He wrote it on the inside. He is the only one who has seen it.} \\
  \textbf{Curse cost:} Opening the chest ticks \textbf{Curse Awakening +3}. The governor's ghost will appear and offer a bargain: the name in exchange for a service. Refusing means the chest locks forever.

\item \textbf{Mine-Share} \hfill \diamondcard{10} \textit{[RESOURCE] [ALLIANCE]} \\
  In the commune; they fight like they mean it. \\
  \textit{The share is a pickaxe. The commune will lend you a miner---or a grave. The mine was opened by the empire's first surveyors. They found iron. They also found something else. They did not name it.} \\
  \textbf{Curse cost:} The miner will work for you, but the mine's whispers will follow. Once per session, the GM may ask you to name something you found in the dark. If you lie, the miner dies.

\item \textbf{Blood-Peace Charter} \hfill \diamondcard{J} \textit{[SOCIAL] [TRUCE]} \\
  Suspend a feud long enough to move your wagons. \\
  \textit{The charter is a white ribbon. Blood will not be spilled while it flies---but blood will remember. The ribbon was woven by a hedge-witch. She used thread from a tax collector's uniform. The collector is dead. The thread is not.} \\
  \textbf{Curse cost:} The truce lasts one scene. After that, the feud returns with interest — both sides will demand a wergild from you. Pay 2 Supply or start a \textbf{Double Feud [6]}.

\item \textbf{Marriage Proxy} \hfill \diamondcard{Q} \textit{[SOCIAL] [ALLIANCE]} \\
  From the Margravine; bind a hill-king to your cause. \\
  \textit{The proxy is a wedding veil. The veil was worn by a ghost. The ghost was the Margravine's grandmother. She married an imperial legate. The legate never came home.} \\
  \textbf{Curse cost:} Using the proxy binds the hill-king to you — but the ghost will appear at your own wedding (or any celebration). She will ask for a dance. Refusing costs 1 Boon and ticks \textbf{Curse Awakening +1}.

\item \textbf{The Lame King's Traveling Writ} \hfill \diamondcard{K} \textit{[AUTHORITY] [ONCE]} \\
  Troops must make way (for now). \\
  \textit{The writ is a broken scepter. The Lame King's seal is a footprint. The footprint was pressed into wax by his grandfather. The grandfather was a free company captain. He could not write. He used his boot.} \\
  \textbf{Curse cost:} The writ works only once. After that, the Lame King will send debtors to collect the scepter. If you refuse, start a \textbf{King's Hunt [6]}.

\item \textbf{Curse-Redemption Rite (Temporary)} \hfill \diamondcard{A} \textit{[RITUAL] [ONCE]} \\
  For one night, no crossroads can hold you. \\
  \textit{The rite is a whisper. The whisper is a name you have never spoken. The name is the Curse's. It does not want to be spoken. It does not want to be silent. It wants to be paid.} \\
  \textbf{Curse cost:} Using the rite clears all \textbf{Curse Awakening} segments for one night. The next morning, the Curse returns with double the segments. You cannot use the rite again in the same province.
\end{enumerate}

\section*{Quick use notes}
\label{sec:acasia-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (papers/claims/safe-conducts) that change position rather than call for a roll.
\item If any A appears, echo plague and curse motifs---crossroads that turn back on themselves, bells tolling for no funeral, banners blackened by mildew, the ghost of the empire in every milestone.
\item Track the \textbf{Curse Awakening [10]} timer as a campaign-wide pressure. Use the \textbf{Curse Stirrs} table when it reaches 8.
\end{itemize}

\subsection*{Additional Features}
\begin{itemize}
\item \textbf{Feud as Currency:} Any insult, accident, or rumor can birth a feud; the GM may replace an SB result with a feud-token tied to a petty lord or house. Feud-tokens act as reusable complications: once per scene, the GM can spend a feud-token to have a rival appear, a bridge close, or a village refuse aid.
\item \textbf{Rot \& Tithe:} Each market or harvest scene carries a risk of rot---add a pressure if food or coin are exchanged without a blessing or a seal. A failed roll may double the tithe or spoil stored supplies.
\item \textbf{The Curse Remembers:} When roads or negotiations stall, the GM may quietly repeat imagery---crossroads, whispers, mildew---to remind players that Acasia itself resists resolution. If the party ignores three such omens, the \emph{Curse Stirrs} complication triggers automatically.
\item \textbf{Empire's Ghost:} Any ancient imperial structure (milestone, bridge, fortress) may still be haunted by the echo of its original purpose. A character who makes an offering (a coin, a salute, a whispered oath) may gain +1 die on a single roll within sight of that structure. The offering must be sincere. The empire does not accept bribes. It accepts memory.
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Bitter Root:} Daughters of the Cord run a hospice in a burned abbey. They ask no payment---but they cut threads. A renegade Sister hides among the charcoalers, trying to untie her cord from a Rothari death warrant. She says the empire's last weight is still unpaid.
  \item \textbf{The Threshold Compendium:} The Pale Causeway is a Class II Anchor site---the ghosts of workers who died building it still carry stones. Saikou Ira would leave a silver coin in every crack. The workers do not spend the coins. They stack them. They are building something.
  \item \textbf{Malachai, the Chained Angel:} A silver coin that never spends circulates among the toll-collectors. Take it, and your passage is always free---but the coin will taste of blood every time the Causeway floods. The coin was minted in the empire's last year. The emperor's face is worn smooth.
  \item \textbf{The Grey Wanderer's Grimoire:} Acasian psions are called ``Feud-Readers.'' They taste the weight of old grudges in the water. The Margravine employs two---and has never trusted a settlement they didn't whisper. They say the Curse is not a thing. It is a conversation. The empire is still talking.
  \item \textbf{The Ninth:} In any Acasian scene, the ninth toll paid at a bridge is never the last. The wise pay eight and barter for the ninth. The ninth toll is collected by a child with no name. The child is the Curse's accountant.
\end{itemize}

\subsection*{Acasian Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Milestone:} Crumble a crust at a broken milestone. Ignore the first \emph{Peat-Fog} penalty.
  \item \textbf{Knot the Contract:} Tie a single knot in a writ's ribbon. Cancel \emph{Bridge Feud} or take –1 die on your next negotiation (your choice).
  \item \textbf{Name to the Blight:} Whisper a dead tax collector's name into a grain sack. Gain +1 Effect when avoiding tithes this scene, but mark \emph{Obligation} to the Margravine.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Decline \& Patronage}]
Patronage here is as brittle as candle wax. The Rothari Clan holds their marches by fear and iron law---patronage from them is real, but cruel, demanding service in blood or levy. Elsewhere, noble names are little more than titles printed on parchment and sold like promises. The Margravine keeps a tally of feuds, binding bloodlines to her feast days, while mercenary captains trade loyalty for coin, switching colors with each bell.

To seek a patron in Acasia is to choose which hunger you will feed: the Rothari's iron teeth, a Free Company's purse, or the hollow blessing of a saint who may never have lived. Yet even false crowns cast shadows long enough to shelter those desperate enough to kneel. The empire is dead. Its weight is not. And the Curse is patient.
\end{tcolorbox}

\section*{Thematic SB Spend Table}
\label{sec:acasia-sb}

\subsection*{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{tithe-collectors or petty lords}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{Free Company patrols or village watchmen}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{feud-trackers or hedge-witches}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable road conditions}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\subsection*{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{River reeve or Free Company captain} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{fog or bridge closure}.
\item \textbf{Foe Appears:} A \textbf{rival bloodline or toll-lord's enforcer} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\subsection*{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{feud-kin, Free Company deserters, or Margravine's guards} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes - \textbf{contract voided/levy doubled/curse triggered}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\subsection*{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Margravine, Lame King, or Cursed Child} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically - \textbf{fog thickens/bridge falls/curse loops}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice - \textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\subsection*{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Acasia (Feud \& Tithe):} Toll collectors demand double, bloodlines take offense, wagons are seized as ``payment.''
\item \textbf{Acasia (Rot \& Famine):} Grain blights spread, wells run dry, abandoned villages hold only ghosts.
\item \textbf{Acasia (Curse \& Crossroads):} Roads loop back, milestones point the wrong way, the same bridge appears twice.
\item \textbf{Acasia (Empire's Ghost):} A phantom legionary marches past, an imperial seal glows on a broken stone, a tax collector's ghost demands arrears.
\end{itemize}

\subsection*{Boss Archetypes in Acasia}
\label{sec:acasia-bosses}
\begin{itemize}
\item \textbf{The False King (Feud/Authority)} --- A petty lord or pretender who crowns themselves on the ruins of empire. \emph{Phases}: Levy Demanded $\rightarrow$ Feud Declared $\rightarrow$ Hasty Coronation. \emph{Cracks}: a \textbf{Blood-Peace Charter} suspends his authority; a \textbf{Marriage Proxy} binds him to a rival.
\item \textbf{The Witch of the Blackwoods (Witchcraft/Curse)} --- A hedge-witch, matriarch, or revenant who manipulates feuds and curses. \emph{Phases}: Tithe of Fear $\rightarrow$ Curse of the Crossroads $\rightarrow$ Hollow Wedding. \emph{Cracks}: a \textbf{Curse-Redemption Rite} breaks her hold; a \textbf{Mine-Share} gives the commune courage to defy her.
\item \textbf{The Tithe Collector (Greed/Extortion)} --- A bureaucrat-priest with writs, escorts, and the ability to summon feuds or famine with a single decree. \emph{Phases}: Audit of Souls $\rightarrow$ Seizure of Grain $\rightarrow$ The Reckoning. \emph{Cracks}: a \textbf{Toll-Exemption Plaque} voids his claim; a \textbf{Wine-Right} buys off his debtors.
\item \textbf{The Rotting Child (Mystery/Curse)} --- A folkloric horror tied to Acasia's plague of false crossroads. \emph{Phases}: Laughter in the Fog $\rightarrow$ The Bridge That Loops $\rightarrow$ The Feast of Ends. \emph{Cracks}: a \textbf{Tithe-Remission Writ} fed to the fog; a \textbf{Monastery Letter} read aloud at the crossroads.
\item \textbf{The Imperial Ghost (Undead/Precedent)} --- The shade of the last imperial governor, still trying to collect the taxes. \emph{Phases}: Audit of the Damned $\rightarrow$ Summon the Debtors $\rightarrow$ The Emperor's Peace. \emph{Cracks}: produce the original imperial charter (in the Sealed Dowry Chest), pay a wergild in the empire's own debased coinage, or convince the ghost that the empire is truly dead.
\end{itemize}

\subsection*{Encounter Seeds (Quick Hooks)}
\begin{itemize}
\item \textbf{The Dowry That Walked:} A Blackwood matriarch's dowry chest has been stolen. The chest contains a claim to a bridge that never floods---and a curse that the Margravine wants broken. The chest was last seen in the hands of a free company captain who claims he is the heir to the empire.
\item \textbf{The Lame King's Feast:} The Lame King invites the party to a feast. The feast is a trial. The food is memory. The wine is lies. The king wants to know who killed the last imperial surveyor. The surveyor is sitting at the table.
\item \textbf{The Curse of the Pale Causeway:} The Causeway is repeating. Every crossing brings the party back to the same inn. The innkeeper is missing a finger. She says she gave it to the bridge. The bridge wants more. The bridge remembers the empire's last toll.
\end{itemize}

% Index entries
\index{Acasia|see{Broken Marches}}
\index{Theme generator}
\index{Broken Marches generator}
\index{Places!Acasia}
\index{People!Acasia}
\index{Complications!Acasia}
\index{Rewards!Acasia}
\index{Utaran Empire!Acasian province}
\index{Curse of Acasia}
\index{Margravine of the Broken March}
\index{Lame King}
\index{Pale Causeway}
\index{Silkstrand}
\index{Imperial surveyors}
\index{Free companies}
\index{Hedge-witches}
\index{Blood feuds}
\index{The Ninth in Acasia}