\section{Aelinnel (Gnomes) --- ``Stone, Bough, and Bright Things''}
\label{chap:aelinnel-gnomes}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{Count the steps, weigh the words, and never trust a single telling.}

\begin{quote}
  \textbf{Lady of Thorns (Fae Sovereign):}  
  ``The law here flows like water over stone – it finds its own path, and woe to those who try to dam it with paper and pretense. The earth itself will testify against the unjust. But the earth also lies – politely, and only when the moon is full.''
\end{quote}

\begin{quote}
  \textbf{Charcoal-burner (Commoner):}  
  ``Our kilns tell us more than your books ever could. When the smoke curls left, the tide's with you. When it splits in three, best stay by your own hearth till it burns straight again. And never trust a court that speaks in only one voice – the silence is always the second witness.''
\end{quote}

\textbf{Genre / Mood:} Dark faerie tale, moonlit groves, polite iron, and the uncanny weight of a promise unsaid.

\textbf{Starting Location:} A tide-shed at dawn, fog curling over anchored barges, gulls crying omens over the river's black glass. A gnomish reeve offers passage on a sealed barge: ``The Green Gate opens at neap tide. What we trade there, the Dolmis will not name. And if you hear someone counting your footsteps, do not turn around – they are only making sure you still have them.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Stone, Bough, and Bright Things}]
The Aelinnel do not build cities. They \emph{grow} them – from granite and hawthorn, from tide-polished bone and the shadow of a laugh. Their halls thread through sea-cliffs and root-galleries, their roads are counted by antler-posts that move at night, and their doors open only when you have forgotten why you came. Here, promises have \emph{weight} – not the weight of ink on parchment, but the weight of a stone dropped into a still pool. The ripples go on forever.

To the north, the Valewood breathes; to the south, the downs of Amedell where the Aelaerem tie red threads to keep the Hollow at bay. The Green Gate sometimes opens onto halfling barrows – a fact neither people speaks of aloud, because speaking aloud gives the gate permission to move.

No single crown rules the gnomish lands. Dozens of courts, seasonal pacts, and garden societies weave a shifting lattice of loyalty. Some Aelinnel keep the old fae treaties, speaking in two voices and never carrying naked iron; others barter in the Green Market with memories instead of coin, treating courtesy as a kind of lockpicking. Still others vanish into the deep woods, listening to stag-roads where the trees whisper gossip about the living. The gnomish heart is divided – and that division is the only reason the fae haven't eaten them whole.

Craft and courtesy govern every traffic between mortal keeps and the courts under the hill. Gnomish stone-singers hum bridges into alignment, their melodies keeping the Green Gate at bay. Forester-wardens hammer copper nails into thresholds because iron, they say, \emph{startles the truth}. In markets under living roofs, a sentence can be folded like a napkin, and a memory can be wrapped for travel – but never opened before the second sunrise. Paths sometimes shorten for those who keep good count, and lengthen for those who sneer at patterns. The fae are never rude – but they are never, ever \emph{safe}.
\end{tcolorbox}

\subsection*{At a Glance}
\begin{itemize}[leftmargin=*]
    \item \textbf{Atmosphere:} Moonlit groves, polite iron, and the uncanny weight of a promise you don't remember making.
    \item \textbf{Key Conflict:} What is said versus what is meant; fae courtesy versus mortal bluntness; the ninth step that is never there.
    \item \textbf{Notable Threats:} Glamour fog that folds your thoughts, iron offence (the wrong nail at the wrong door), geas-catch (a phrase that becomes a leash), the Green Gate yawning at the wrong hour.
    \item \textbf{Common Treasures:} Hazel tokens that remember favours, tide-path keys that work only at ebb, oath-beads that glow when a promise shifts, and the Wild Hunt's clemency – which expires at dawn.
\end{itemize}

\subsection*{Prominent NPCs of Aelinnel}
\label{sec:aelinnel-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Thorn & Lady of Thorns, Sovereign of the Hawthorn Court & Remind mortals that courtesy is a leash, not a gift & She never repeats herself. Her silences are verdicts – and her verdicts are always polite. \\
\hline
Caelum Viridant & Green Knight, Warden of the Stag-Roads & Test the worthy and trap the careless & His armour is woven from living boughs that bleed when he lies. He has never bled. \\
\hline
Eilidh Stone-Prince & Mortal ruler of the Hall of Aelinnel & Keep the old pacts indexed and the fae at arm's length & She speaks only in conditional tense. No one has heard a declarative sentence from her lips in seven years – not even her cat. \\
\hline
Riven & Huntsman of the Moonlit Ride & Chase down obligations and those who flee the Green Gate & His horn is carved from a fossilised scream. He only blows it when a promise is nine days old – and nine days old is when the Hollow starts counting. \\
\hline
Tallow & Charcoal-burner (commoner exemplar) & Read the smoke and warn the village before the fae come & He has never lost a sheep to the Hollow. He has lost three helpers who did not listen. Their shadows still work his kilns. \\
\hline
Sable & Green-Market Broker & Trade truths for trinkets, and never lose the margin & Her receipts are written on leaf-vein paper. They curl when the bargain is unbalanced – and also when it rains, which is often, which is why she keeps a stone in her pocket. \\
\hline
\end{longtable}

\subsection*{Names of Aelinnel}
\label{sec:aelinnel-names}

Aelinnel names are built from soft vowels (ae, ei, ui, oi) and liquid consonants (l, r, n, m) with occasional sharp endings (-k, -t, -x). Surnames are often taken from a notable grove, a tide-mark, or an ancestor's defining courtesy. They evoke the sound of water over stone and wind through hawthorn – and the silence that follows when both stop.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Caelum & Eilidh & Thornweave & \textit{the Green-Wise} \\
\hline
Riven & Sable & Tidespeak & \textit{Two-Tongue} \\
\hline
Tallow & Merrow & Evenstep & \textit{the Unseen Guest} \\
\hline
Fennel & Briar & Hawthorn & \textit{Barrow-Kept} \\
\hline
Osprey & Vesper & Silverbough & \textit{the Uncounted} \\
\hline
Linden & Ash & Foxtail & \textit{Ash-Finger} \\
\hline
\end{longtable}

\subsection*{The Three Courts: Aelinnel Alignments}
Aelinnel society is not a single crown but a lattice of courts, garden societies, and seasonal pacts. Three broad stances shape their dealings with outsiders and with the fae. Most gnomes shift between these postures depending on the season, the tide, and whether the mushrooms are singing.

\begin{itemize}[leftmargin=*]
    \item \textbf{Thorn–Courteous (Fae–Lean)} – These gnomes keep the old pacts with the Lady of Thorns and the Green Knight. They speak in two voices, never carry naked iron, and treat every threshold as a negotiation. They are the ones who remind visitors to count their steps and pour no ninth cup – and who smile when you ask why.  
    \emph{Clocks: \textsc{Hawthorn Patience} [4], \textsc{Courtesy's Leash} [6].}

    \item \textbf{Market–Square (Trade–Lean)} – These gnomes prioritise the Green Market, where memories are sold by the weight and names are traded like buttons. They are pragmatic, will use iron if it is gilded, and view fae courtesies as bargaining chips rather than sacred laws. They are also the most likely to owe a favour to a talking badger.  
    \emph{Clocks: \textsc{Market Moot} [4], \textsc{The Unbalanced Scale} [6].}

    \item \textbf{Forester–Trusted (Wild–Lean)} – These gnomes live among the stag-roads and root-galleries, closer to the stone-singers and the tree-wardens. They speak the old names of tides and groves, and they are the first to notice when the thorn blight spreads or the moonwell stains a name wrong. They also know which mushrooms to eat and which mushrooms eat you back.  
    \emph{Clocks: \textsc{Root Calling} [4], \textsc{Wild Hunt Interest} [6].}
\end{itemize}

Most Aelinnel characters and settlements fall somewhere between these poles, and the same gnome might wear a different mask when trading at the Green Market than when walking the stag-roads. The masks, incidentally, are not optional.

\section*{Spades --- Places (stone spires, sea-rock, deep woods)}
\label{sec:aelinnel-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Tide-rift steps} – Cut into black rock; limpets and old votive nails keep count for those who forget. The ninth step is underwater – and it still counts. `[THRESHOLD]`
\item \textbf{Moonwell basin} – A granite bowl; coins turn green in a week, names in a season, and memories in a year. The water never reflects the same face twice. `[MAGIC]`
\item \textbf{Dolmen stair} – Up a ridge; each capstone rings a different interval if you tap it true. The ninth stone is silent – it only rings when someone lies. `[MUSIC]`
\item \textbf{Charcoal coppice} – Under witch-hazel; a neat ash-heap hides a door that opens on even knocks only. Knock an odd number, and the door will open – but not where you left it. `[SECRET]`
\item \textbf{Basalt organ cliffs} – Sea-caves breathe like slow bellows; the wind prefers prime numbers. When it plays a ninth, the cliffs hum back. `[NATURE]`
\item \textbf{Stag Road} – Game-trail marked by antler-posts; hooves know it best, maps accept it grudgingly. The posts move at night – not to confuse, but to keep the road honest. `[TRAVEL]`
\item \textbf{Quartz spring} – Sand shines like ground stars; drinkers tell the same story with better structure. The ninth sip tastes of copper – and a promise you haven't made yet. `[HEALING]`
\item \textbf{Elf-causey} – Pale flags through a marsh; visible at dawn and dusk and whenever someone is counting aloud. Count to nine, and the flags turn to water. `[FAE]`
\item \textbf{Root gallery} – Under an oak hill; lanterns hang from living bark – two lit, one listening. The listening lantern is for the Hollow. It never lights. `[UNDERGROUND]`
\item[J] \textbf{Barrow gallery} – Stone chimes; a cold draft from below plays scales no throat can sing. The ninth note is always a sigh. `[DEATH]`
\item[Q] \textbf{Thorn Court ring} – Hawthorns trained into arches; petals fall like knife-points and settle into proofs. Step into the centre, and the thorns will ask you a riddle. Answer wrong, and you become a hedge. `[FAE]` `[DANGER]`
\item[K] \textbf{Hall of Aelinnel} – A timber keep threaded between standing stones; floors level themselves for guests. The ninth stone in the floor is always warm – it remembers the last person who lied. `[CIVILISATION]`
\item[A] \textbf{The Green Gate} – In a sea-arch; at certain tides it opens to a forest not on any map, and expects exact change. It does not accept coin. It accepts \emph{explanations}. `[GATE]`
\end{enumerate}

\section*{Hearts --- People \& Factions (keepers, courts, craftsmen, hunters)}
\label{sec:aelinnel-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Goat-herd of the stone edges} – Measures danger in hoof-widths; never wrong twice. The third time, she doesn't measure – she runs. `[SURVIVAL]`
\item \textbf{Charcoal-burner clan} – Read draught and omen by smoke; their kilns hum in thirds. When the hum stops, someone has asked a question they shouldn't have. `[DIVINATION]`
\item \textbf{Hedge-witch} – A pocket of thorns and prices stated in \emph{unlesses}. Her fees are always conditional: ``I will heal your child, \emph{unless} the moon is full, in which case I will only listen.'' `[MAGIC]` `[SOCIAL]`
\item \textbf{Stone-singer (quarryman)} – Hum finds a seam and a secret; walls relax when they pass. The ninth hum is for doors that do not exist – yet. `[CRAFT]`
\item \textbf{Forester-warden} – Copper nails and a polite dislike of iron; paths obey their hammer. They never cut a branch – they ask permission, and sometimes the branch says no. `[NATURE]`
\item \textbf{Reed-net fisher} – Knows when the selkies listen; mends nets to lullabies in 5s and 8s. The ninth verse is always for the drowned. `[TRADE]`
\item \textbf{Oath-carver} – Sets promises in quartz so they can be seen; breaks none, bends many. A broken quartz-oath turns to sand – and the sand remembers. `[LAW]` `[OATH]`
\item \textbf{Way-keeper of the Stag Road} – Never lost, often followed, sometimes found ahead. When she stops at a crossroads, she always turns her back to the ninth path. `[GUIDE]`
\item \textbf{Green-market broker} – Trading truths for trinkets and back again; receipts in leaf-vein script. The truth she sells is always true – but never the truth you asked for. `[COMMERCE]`
\item[J] \textbf{Green Knight} – Antler helm, mirror-bright blade, courtesy old as frost; keeps a tally of saved insults. He never unsheathes his sword – he only shows it, and that is usually enough. `[FAE]` `[DUEL]`
\item[Q] \textbf{Lady of Thorns} – Sovereign of a hawthorn court; smiles like a snare and forgives in exact measures. Her forgiveness is always exactly one thorn's weight – no more, no less. `[FAE]` `[LAW]`
\item[K] \textbf{Stone Prince of Aelinnel} – Mortal crown with fae obligations properly indexed and current. He is the only gnome who has never lied – because he never speaks in declaratives. `[AUTHORITY]`
\item[A] \textbf{Huntsman of the Moonlit Ride} – His horn turns shortcuts into roads, and obligations into destinations. When he blows the ninth note, the hunt ends – but the hunted never comes back. `[HUNT]`
\end{enumerate}

\section*{Clubs --- Complications/Threats (glamour, geasa, weather, neighbours)}
\label{sec:aelinnel-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Glamour fog} – Time drifts, footfalls don't match, conversations arrive neatly folded. Open a sentence, and you may find last Tuesday inside. `[ILLUSION]`
\item \textbf{Iron offence} – Someone brought the wrong nails to the right grove; paths take it personally. The trees do not speak – they \emph{remember}. `[SOCIAL]` `[WEIGHT]`
\item \textbf{Root-slide} – Pulls a trail two ridges away while you blink; your footprints stay where they were. The footprints will wait for you – for nine days. After that, they walk home without you. `[TRAVEL]`
\item \textbf{Wrong tide at the sea-arch} – Boats rise where arithmetic says no and physics excuses itself. The ninth wave is always the wrong one – it goes sideways. `[WEATHER]`
\item \textbf{Spoken geas} – Catches on an unlucky word; the task names itself and refuses synonyms. You cannot say ``fetch'' when you meant ``steal'' – the geas knows the difference. `[MAGIC]` `[OATH]`
\item \textbf{Green Market price} – Payment demanded in names and memories; change returned in riddles. The riddles are always about thresholds. The answers are always doors. `[COMMERCE]` `[WEIGHT]`
\item \textbf{Stone-wight stirs} – In a barrow; echoes hate company and file complaints as drafts. The complaints are written in a language that only the dead speak – and they are \emph{very} articulate. `[DEATH]`
\item \textbf{Lost day} – The sun miscounts; your dawn arrives at supper and demands bread. The bread must be from a loaf you have not yet baked. `[TIME]`
\item \textbf{Thorn blight} – Crawls across orchards; pruners bleed stories that grow where they drip. The stories are always about a door you left open. `[CURSE]`
\item[J] \textbf{Stag horn sounds} – Everyone owes the chase – especially you, especially now. The horn does not say who is hunting. It only says that the hunt has begun. `[HUNT]` `[WEIGHT]`
\item[Q] \textbf{Thorn Court tithe} – Levied on the Hall; arrests wear flowers and quote etiquette. The charge is always \emph{insufficient courtesy}. The penalty is always \emph{a lesson}. `[LAW]` `[SOCIAL]`
\item[K] \textbf{Muster of the Bough} – Forest banners rise; travel becomes trespass by default. The banners were trees yesterday. They are still adjusting. `[CONFLICT]`
\item[A] \textbf{Green Gate yawns} – At the wrong hour; roads rewire across your path with excellent logic. The logic is always sound. The destination is always somewhere else. `[GATE]` `[MYSTERY]`
\end{enumerate}

\section*{Diamonds --- Treasures \& Discoveries (charms, passes, host-rights, exceptions)}
\label{sec:aelinnel-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Hazel token} – Step through a hedge uncut, once; the scratch you did not get will \emph{remember} you kindly. The token is a nut. Crack it open, and the hedge will part – but it will also ask a question. `[PROTECTION]`
\item \textbf{Salt license} – Carry and use ward-salt where it's discouraged; doors frown and open anyway. The license is a pinch of salt in a silk pouch. The silk was a promise. `[UTILITY]`
\item \textbf{Dolmen carving-right} – Inscribe a name or claim on a speaking stone; arguments quiet while you carve. The stone will remember your chisel's touch. It will also remember what you were thinking. `[AUTHORITY]`
\item \textbf{Guest-bough} – From the Thorn Court; one night's safe table among thorns; knives behave. The bough is a twig with three thorns. Break one thorn for each courtesy you ask. When all three are broken, the table becomes a door. `[HOSPITALITY]`
\item \textbf{Quarry allotment} – Of clear quartz; subtle doors open to its bearer if asked in the right tense. The right tense is not past or future – it is \emph{conditional}. `[ACCESS]`
\item \textbf{Green Market voucher} – Buy a truth at face value (today only); tomorrow costs interest. The interest is always a memory – and the voucher chooses which one. `[COMMERCE]`
\item \textbf{Oath-bead} – A spoken promise warms the bead when kept, chills when not, glows if misunderstood. A glowing bead is not a broken promise – it is a promise that has learned to speak for itself. `[OATH]`
\item \textbf{Tide-path key} – Lawful crossing of a named sea-cave at neap; the cave will wait. The key is a piece of dried seaweed. It will dissolve after one use – but the cave will remember your face. `[TRAVEL]`
\item \textbf{Forest truce} – Hunt, cut, or travel through a warded copse without offence; the birds file no reports. The truce is a single leaf. Carry it in your mouth, and the forest will not see you – but the Hollow will. `[PERMISSION]`
\item[J] \textbf{Green Knight escort} – From ridge to gate; watchers bow, paths part, insults lodge elsewhere. The Knight does not speak during the escort. The silence is the escort. `[PROTECTION]`
\item[Q] \textbf{Private audience} – With the Lady of Thorns; a whisper exits as policy with petals attached. The petals are sharp. Do not handle them carelessly – they cut what they agree with. `[AUTHORITY]`
\item[K] \textbf{Stone Prince's seal} – Levies defer, foresters guide, scribes stop arguing and fetch tea. The seal is a pebble. It only works when you are not looking at it. `[AUTHORITY]`
\item[A] \textbf{Wild Hunt clemency} – Ride under the horn for one night; no hound will take you, no obligation will sleep. The clemency is a silver bell. Ring it, and the Hunt will pass you by – but the Hollow will ride behind you until dawn. `[PROTECTION]`
\end{enumerate}

\section*{Quick use notes}
\label{sec:aelinnel-quick-use}
\begin{itemize}
\item Draw until all four suits appear: \textbf{Spade} = place, \textbf{Heart} = actor, \textbf{Club} = pressure, \textbf{Diamond} = treasure. Highest rank sets the main Clock (2–5 → 4, 6–10 → 6, J/Q/K → 8, A → 10).
\item Treat \textbf{Diamonds} as codified outcomes (charms/passes/host-rights) that \emph{change position} rather than call for a roll.
\item If any \textbf{A} appears, echo dark fae motifs: antlers in shadow, petals that cut, a tide that breathes, and a shortcut that insists on proper payment – and always, always a ninth step that is not there.
\end{itemize}

\subsection*{Additional Features}
\begin{itemize}
\item \textbf{Two-Voice Courtesy:} Once per scene, a character who states both what they \emph{say} and what they \emph{mean} (in or out of character) may shift \emph{Position} one step safer on a social roll against a fae or fae-touched being. The GM may demand a small token (a copper coin, a knotted thread, a single hair) as earnest – and the token will never be returned.
\item \textbf{Copper over Iron:} When the party visibly favours copper or brass tools over iron in fae-facing scenes, gain +1 assist die from locals (or avoid a –1 penalty) for acts of courtesy, passage, or petition. Iron is not forbidden – it is simply \emph{impolite}, and impoliteness is remembered. Copper, by contrast, is \emph{interesting}. The fae like interesting.
\item \textbf{The Unsaid Clause:} In any bargain or promise, if the players explicitly leave one term unspoken (but agreed upon by all at the table), the GM may bank 1 SB to introduce a complication later where that omission becomes the hinge of the deal. The fae love what is not said – they build entire courts out of it.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Thorned Lady} – The Lady of Thorns (fae sovereign). \textit{Courtesy is a leash; mercy is a snare; and the ninth cup is always for the guest you cannot see.}
    \item \textbf{The Green Knight} – The Sealed Gate (threshold guardian). \textit{Prove your worth, or be turned aside – but worth is not what you carry. It is what you leave behind.}
    \item \textbf{The Still Count} – Mykkiel (divine arithmetician). \textit{Every miscount is a sin; every correction is a prayer; and the ninth number is always a whisper.}
    \item \textbf{The Root-Speaker} – Grimmir (wild patron). \textit{Stone sings, wood remembers, the grove witnesses – and the ninth root is always listening.}
  \end{itemize}
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Courtesies of the Green Gate}]
\begin{itemize}
\item \textbf{Never bring naked iron} past a hawthorn arch unless it is gilded or named. Copper is polite, silver is opinionated, cold iron is an accusation. The hawthorn will remember which you carried – and will tell the gate.
\item \textbf{Always return what points the way.} Way-cords, antler-posts, chalk. Borrow the path; do not keep it. The fae count what you take. They also count what you leave.
\item \textbf{Speak obligations in the daylight.} Promises made under lantern-boughs are heard by leaves as well as by law. The Green Gate listens – and leaves have long memories.
\item \textbf{The ninth cup is never poured.} Leave it empty for the uninvited guest. If you pour it, drink it yourself – and hope the guest does not arrive. The guest is always polite. The guest is always hungry.
\end{itemize}
\end{tcolorbox}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Bitter Root:} Daughters of the Cord sometimes serve as oath-carvers, binding promises with red thread instead of quartz. A renegade Sister hides in the Root Gallery, trying to untie a cord tied by the Lady of Thorns herself. The cord has nine knots – and the ninth knot is a smile.
  \item \textbf{Malachai, the Chained Angel:} A silver coin that never spends appears in the moonwell. Take it, and the water will show you a perfect reflection – but the coin will reappear in the well every new moon, each time a little warmer. The ninth time, the water shows you a door.
  \item \textbf{The Grey Wanderer's Grimoire:} Aelinnel psions are called “Proof-Readers”. They hear the weight of hidden clauses and can detect an omission in any bargain by the way the air tastes of copper. The ninth omission always tastes of honey – and that is how you know it is a trap.
  \item \textbf{The Threshold Compendium:} The Green Gate is a Class III Procession – the gate itself walks the coast once a season, and those who cross it must answer a riddle about obligations or be pulled into the Hollow. The riddle is always the same: “What do you owe to the door that opens?” The answer is never the same twice.
  \item \textbf{The Ninth:} In any Aelinnel scene, the ninth word spoken in a bargain becomes legally binding – even if the speaker did not intend it. The wise count to eight and then gesture. The ninth word, if spoken, is always heard by the gate. The gate always laughs.
\end{itemize}

\subsection*{Regional Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Gate:} Crumble a crust before the Green Gate. Ignore the first \emph{Glamour fog} penalty. The crust must be from a loaf you baked yourself – or from a loaf you stole from someone who baked it with love. `[SUPERSTITION]`
  \item \textbf{Knot the Cord:} Tie a single knot in a way-cord. Cancel \emph{Spoken geas} or take –1 die on your next promise (your choice). The knot must be tied while humming. The hum does not need a tune – only a breath. `[SUPERSTITION]`
  \item \textbf{Name to the Stone:} Whisper a dead oath-carver's name into a standing stone. Gain +1 Effect when negotiating with fae this scene, but mark \emph{Obligation} to the Stone Prince. The stone will whisper back – but only once. Listen carefully. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table (Aelinnel)}
\label{sec:aelinnel-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{thorn-courtiers or forest-wardens}. The attention is always polite – and always dangerous.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{stone-singers or green-market traders}. The stone-singers do not move – they only listen. The traders always move – they only pretend to listen.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{stag-road trackers or oath-carvers}. The trackers are deer. The deer are not deer.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}. The tide is never favorable – it is only patient.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}. The clock ticks in eights. The ninth tick is a silence.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Green Knight or Lady of Thorns' herald} becomes aware and begins responding. The herald is a rook. The rook speaks in couplets. The couplets are always warnings.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{glamour fog or root-slide}. The fog is not fog – it is folded time. The root-slide is not roots – it is a memory of a landslide.
\item \textbf{Foe Appears:} A \textbf{stone-wight or thorn blight spirit} arrives on scene. The stone-wight does not attack – it asks questions. The thorn blight does not bleed – it grows.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}. The equipment is not broken – it has simply forgotten its purpose. Remind it with a story.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome. The obstacle is a door. The door has no handle – only a riddle.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{fae courtiers, forester-wardens, or stag-horn riders} arrive. The courtiers are polite. The wardens are patient. The riders are neither.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable. The tool is not broken – it has chosen to rest. Apologise to it.
\item \textbf{Major Twist:} The situation fundamentally changes – \textbf{geas triggers / green gate opens / iron offence declared}. The geas was always there – you just did not hear it. The gate was always open – you just did not see it. The offence was always waiting – you just did not apologise.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment. The clock ticks in eights. The ninth tick is a silence – and silence is always a verdict.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction. The condition is always \emph{courteous}. The courteous are always tired.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect. The trap was not set – it was \emph{suggested}. The suggestion was polite. The effect is not.
\item \textbf{Authority Arrival:} \textbf{Lady of Thorns, Stone Prince, or Huntsman of the Moonlit Ride} intervenes. The Lady never arrives – she was always there. The Stone Prince never leaves – he was always waiting. The Huntsman never stops – he was always riding.
\item \textbf{Scene Shift:} The environment changes dramatically – \textbf{tide turns / green gate yawns / thorn blight spreads}. The tide turns sideways. The gate yawns – and it has teeth. The blight spreads in eights – the ninth is a flower.
\item \textbf{Patron Omen:} Divine/arcane forces take notice – \textbf{omen appears / blessing lost / curse manifests}. The omen is a single word. The blessing is a door closing. The curse is a promise kept.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives. The turn was always there – you just did not count the steps.
\end{itemize}

\paragraph{Region-Specific SB Options (Aelinnel)}
\begin{itemize}
\item \textbf{Courtesy \& Geas (1–3 SB):} A hawthorn arch closes behind you; copper tools glow with fae attention; a polite phrase becomes a binding geas. The fae have noted your manners – or your lack thereof. The note is kept in a drawer. The drawer is locked. The key is a word you have not said.
\item \textbf{Glamour (1–2 SB):} Conversations arrive folded and misaddressed; time drifts unexpectedly; your footfalls no longer match your steps. The Green Gate breathes – and when it breathes out, the world folds. When it breathes in, the world waits.
\item \textbf{The Unsaid (1–2 SB):} An omission you did not know you made becomes the price of passage. The Lady of Thorns smiles. You do not know why. The smile is the price.
\end{itemize}

\subsection*{Aelinnel – Tidal Bargains}
\textbf{Starting Location:} A tide-shed at dawn, fog curling over anchored barges, gulls crying omens over the river's black glass.
\begin{quote}
\textit{``In Aelinnel lands, the river speaks with many tongues, and each tide demands a name in return. The ninth tide demands a secret. The secret must be one you did not know you had.''}
\end{quote}
\textbf{In...} a tide-shed at dawn, a gnomish reeve offers passage on a sealed barge. ``The Green Gate opens at neap tide. What we trade there, the Dolmis will not name. And if the barge creaks nine times before we cast off, do not board – the tide is lying.''
\paragraph{Regional Motifs} Tide-ledgers, river gods, sealed cargo, name-taboo, the ninth knot that is never tied.
\paragraph{Whispers to Offer} ``Peek in the crate,'' ``Speak the hidden name,'' ``Break quarantine – the seal is only wax.''
\paragraph{External Clocks} Threat: \textit{Green Gate Thirst [6]}; Community: \textit{Dockside Fear [4]}; Anchor: \textit{Barge Oath [4]}; Hidden: \textit{The Ninth Tide [8]}.
\paragraph{Complications} Freshwater turns brackish; the cargo sings; a rival reeve boards with knives; the tide arrives an hour early – or a year late.
\paragraph{Anchors at Risk} A PC's river-kin; the barge's blessing; a knot tied by the wrong hand.
\paragraph{Escalation Beat} The Gate opens early; the tide runs backwards; the ninth bell rings – and no one rang it.
\paragraph{Relief} Offer a river-gift; carry a name written in river-mud across the threshold; spend 2 \Boons{} to seal the leak for a scene – but the leak will remember. It will always remember.

\subsection*{Faction Entanglements}
\begin{itemize}
  \item \textbf{Stone-Singer Conclave:} Sponsor a bridge-chant or way-tuning and bank a \emph{Harmonic Favor} – spend to shift \emph{Position} one step safer on any passage across worked stone or living root. Skip their rites and the first crossing this leg gains a free ``Lock/Barrier'' SB as paths sulk. The sulking paths are the most dangerous – they hold grudges.
  \item \textbf{Forester-Wardens:} Accept copper-nail protocols to cancel the first \emph{Fae Offence} SB in a wood scene. Drive iron stakes or break taboo and mark a 4-segment \emph{Polite Apology} clock you must clear before locals assist you again. The apology must be spoken to a tree. The tree will not answer – but it will remember.
  \item \textbf{Green-Market Brokers:} Submit an itemised truth (what was said / what was meant) to reduce barter DV by 1 for the scene. Try to haggle on single-ledger terms, and the market tags you with \emph{Exact Change Only} – all bargains hereafter demand a tithe of memory or name until appeased. The memory is never yours to give – but the market takes it anyway.
  \item \textbf{Oath-Carvers' Guild:} Have a promise set in quartz to gain a \emph{Visible Oath} token; spend to treat one social test as if aided by a reputable witness. Breaking a quartz-oath spawns a 4-segment \emph{Shattered Nuance} clock that complicates future negotiations. The shards of quartz are sharp – and they remember what you promised.
  \item \textbf{Tide-Reeves:} File a tide-plan before neap to earn a \emph{Tide Window} – ignore one tidal timing penalty this leg. Skipping the ledger adds a free ``Wrong Tide'' SB the first time you touch the Dolmis. The wrong tide is always the ninth tide – and the ninth tide always arrives sideways.
\end{itemize}

\subsection*{Cross-Regional Ties \& Soft Power}
\begin{itemize}
  \item \textbf{Way of Silk Interface:} Kon"reh-counted caravans that adopt Aelinnel step-metrics (even steps, counted halts) gain +1 die to avoid ambushes on the Way this arc. Breaking count invites a free ``Glamour Fog'' SB at the next way-shrine. The fog is not fog – it is the road's memory of a step you did not take.
  \item \textbf{Aeler Standards (Rods \& Wards):} Recognise Aeler measures publicly to reduce dispute DV by 1 with dwarven legates; insist on gnomish dual-ledger and foreign merchants grumble – first mixed-market scene gains an ``Assay Doubt'' SB. The doubt is always polite. The doubt is always wrong.
  \item \textbf{Vhasia/Viterra Complications:} Vhasian hedge-duties (wood rites) and Viterra tithe-scripts (grain flows) cross your ledgers. When either region is ``Stressed,'' add +1 DV to Green-Market exports that travel their roads unless you carry a \emph{Forest Truce} or \emph{Tithe Waiver}. The roads are not roads – they are conversations. Conversations can be interrupted.
  \item \textbf{Valewood Fallout:} On a \emph{Valewood Weather} signal, fae proofs spill downslope; counting etiquette cancels the first navigation SB each scene, but any iron shown gains a free ``Fae Complaint'' SB (even among humans). Aelaerem hospitality rites mitigate this: presenting guest-loaf removes one such complaint per hall. The loaf must be shared – not given, but \emph{shared}. Sharing is a kind of promise.
\end{itemize}

\subsection*{Trade Goods \& Consequences}
\begin{itemize}
  \item \textbf{Quartz \& Proof-Stone:} Donate clear quartz to a shrine to bank a \emph{Seen Promise} – once, treat a murky contract as if properly witnessed. Peddle clouded quartz as ``ritual grade'' and spawn a roaming \emph{Counterfeit Proof} inquiry clock. The counterfeit proof is not false – it is simply \emph{incomplete}. The missing part is always the ninth clause.
  \item \textbf{Copper Etiquette Kits:} Issuing copper over iron in mixed courts grants +1 assist die on petitions; flooding markets with cheap copper pins offends foresters – add a ``Copper Resentment'' tag that raises DV in wood jurisdictions until you make amends. The amends must be made to a tree. The tree will ask a favour. The favour will be a riddle.
  \item \textbf{Name-Script \& Memory Wraps:} Legal export of name-script lowers espionage DV by 1 for allies who carry it. Smuggling wraps risks a \emph{Name-Tax} – first customs scene converts 1 SB into a compelled disclosure. The disclosure is always a name. The name is always wrong – but the tax collector will not know that.
  \item \textbf{Tide Rights \& Barge Seals:} Funding barge-seals creates a \emph{Waiting Cave} – once per arc the cave ``waits'' an extra beat for your crew. Ignoring seal-maintenance flips it: first Dolmis crossing suffers ``Wrong Hour'' SB. The wrong hour is always the ninth hour – and at the ninth hour, the tide remembers your face.
\end{itemize}

\subsection*{Patron Networks (Patterns \& Thresholds)}
\begin{itemize}
  \item \textbf{Sacred Geometry:} Chalk a golden-mean spiral at a ford to reroll one failed prediction/trajectory check this leg; doing so in view of fae courts adds a 2-segment \emph{Proof Envied} clock. The fae do not envy the proof – they envy the chalk. Chalk is temporary. The fae are not.
  \item \textbf{The Sealed Gate:} Hawthorn-gilded door-nails act as single-use stay writs: negate one forced entry or pursuit at a hedge-gate. The nail must be driven with a copper hammer. The hammer's head must be warm. The warmth is the memory of a door closing.
  \item \textbf{Lunara:} Counting under moonlight lets you ask one terse yes/no about a hidden path; on a Miss, secrets spread – mark Exposure +1 in the nearest Green Market. The secret that spreads is never the one you asked about. It is always worse.
  \item \textbf{The Traveler:} Antler-posts double as wayshrines; leave a copper and recite a sequence to finish one travel clock tick for free on forest legs. The sequence is always the same: eight steps, then a pause. The pause is the offering.
\end{itemize}

\subsection*{Field Procedures (Use At the Table)}
\begin{itemize}
  \item \textbf{Two-Voice Talk:} When parley stalls, the GM asks: ``What was said? What was meant?'' If players state both, cancel the first social SB this scene; if they refuse, the next bargain demands collateral (memory/name). The collateral is always returned – but never in the same condition.
  \item \textbf{Count or Be Counted:} At any fae-touched crossing, a player who audibly counts in a fitting sequence (steps, heartbeats, breaths) shifts \emph{Position} one step safer for that crossing. Boasting or mocking numbers adds a free ``Geas Catches'' SB. The geas is always about a door. The door is always behind you.
  \item \textbf{Copper Courtesy Check:} Producing copper/brass tools in a ritual space removes the first \emph{Fae Offence} tag this scene; drawing iron instead escalates the next Complication by +1 Cap. The escalation is always polite. The result is never.
\end{itemize}

\subsection*{Diplomatic Favor \& Reputation Echoes}
\begin{itemize}
  \item \textbf{Hazel Favors (minor resource):} Earned by returning way-cords, restoring antler posts, or paying tide-dues. Spend to downgrade a \emph{Glamour} or \emph{Geas} complication once per leg. The favour is a hazelnut. Crack it open – the kernel is a word. The word is a key.
  \item \textbf{Alignment Reputation:}
    \begin{itemize}
      \item \emph{Thorn-Courteous:} Doors of the hawthorn courts open a step easier; human clerks distrust your clauses (first city office DV +1). The distrust is mutual. The hawthorn courts are not kind – they are \emph{accurate}.
      \item \emph{Market-Square:} Brokers vouch your weights; rival stalls circulate riddled slander – add a trailing \emph{Folded Rumor} clock. The rumour is always folded. Unfold it, and you will find a truth you did not want.
      \item \emph{Forester-Trusted:} Paths shorten for you once per journey; port inspectors assume concealed copper – first harbor search gains a free ``Alarm Raised'' SB unless declared up front. The paths shorten because they are afraid of you. The inspectors are not afraid – they are \emph{curious}. Curiosity is worse.
    \end{itemize}
\end{itemize}

\subsection*{Boss Archetypes in Aelinnel}
\label{sec:aelinnel-bosses}

\paragraph{Court \& Courtesy (Fae Sovereigns)}
\begin{itemize}
  \item \textbf{Lady of Thorns (Etiquette Engine)} – Punishes breaches with precise harms, rewards perfect courtesy with literal boons. \emph{Win by satisfying three courtesies in sequence (no iron, two-voice speech, return the waymark).} The third courtesy is always a riddle. The riddle is always about a door.
  \item \textbf{Green Knight (Trial by Pattern)} – Duel of paths and proofs, not blades. \emph{Win by presenting a higher-count route (Fibonacci, even-steps) or yielding a deserved apology at the right number.} The right number is never nine. The right number is always eight – and a pause.
\end{itemize}

\paragraph{Threshold \& Hunt (Environment that Judges)}
\begin{itemize}
  \item \textbf{Moonlit Ride (Wild Hunt)} – Three phases: \emph{Sounding} (horns), \emph{Coursing} (paths shorten for the bold), \emph{Closing the Ring}. \emph{Win by naming safe-clemency, joining the chase without taking a prey, or blowing the right counter-call (2–3–5).} The counter-call must be a question. The question must be about the Hollow. The Hollow does not answer – but it listens.
  \item \textbf{Green Gate (Living Crossing)} – Demands exact change in truths. \emph{Win by paying a memory that fits the posted proof; forcing passage triggers ``Roads Rewired.''} The posted proof is always incomplete. The missing part is the ninth word.
\end{itemize}

\paragraph{Numbers \& Glamour (Logical Anomalies)}
\begin{itemize}
  \item \textbf{Glamour Fog (Layered Illusion)} – Each layer dispelled by a correct count, courtesy, or copper rite. \emph{Win by completing the \textit{Three Proofs}: count aloud, return a borrowed path token, trade a said/meant receipt.} The third proof is always a promise. The promise is always to a shadow.
  \item \textbf{Lost Day (Time Miscount)} – The sun miscounts hours. \emph{Win by reconciling two ledgers (what happened / what should have), then ringing a prime-number chime.} The prime number is always 2, 3, 5, or 7 – never 9. 9 is not a number in Aelinnel. 9 is a \emph{suggestion}.
\end{itemize}

\paragraph{Market \& Oath (Contract Tricksters)}
\begin{itemize}
  \item \textbf{Green-Market Broker (Truth Arbitrage)} – Buys your meanings, sells your words. \emph{Win by catching them in a one-voice statement or producing an Oath-Bead that chills (misunderstood clause).} The misunderstood clause is always the one you left unsaid. The broker knew you would.
  \item \textbf{Oath-Carver Gone Wrong (Quartz Tyrant)} – Binds towns with over-precise promises. \emph{Win by carving a \emph{Release Unless} into communal quartz at cost of a public truth.} The truth must be one you have never told. The quartz will remember it. The quartz will also share it – politely, and only with the Hollow.
\end{itemize}

\paragraph{Stone \& Memory (Ancestral Hazards)}
\begin{itemize}
  \item \textbf{Stone-Wight Chorus (Echo Jurors)} – Resent footsteps, adore proofs. \emph{Win by tapping the dolmen scale that matches the grievance and returning a taken token to the barrow.} The token is always a memory. The memory is always someone else's. Return it carefully – the dead are particular about provenance.
\end{itemize}

\paragraph{Boss Mechanics (Use 2–3 levers)}
\begin{itemize}
  \item \textbf{Courtesy Track [4]:} Each breach (iron bared, skipping greetings, one-voice lie) ticks it. On fill, boss gains \emph{Polite Punishment}: escalate DV +1 and impose a Geas. The Geas is always about a door. The door is always open – you just cannot see it.
  \item \textbf{Counting Proofs (three steps):} Name a pattern, perform it, return a waymark. Completing all three lowers boss tier for one exchange or ends a phase. The pattern is always a sequence of eights. The ninth step is a pause – and the pause is the proof.
  \item \textbf{Said/Meant Ledgers:} Any bargain scene runs two mini-clocks; if \emph{Said} fills first, glamour favours the boss; if \emph{Meant} fills first, PCs gain a \emph{Visible Oath} token. The token is a piece of quartz. It glows when someone lies. It also glows when someone tells the truth – but the colour is different. The difference is subtle. The fae love subtle.
  \item \textbf{Iron Offence Meter [0–3]:} Showing iron, striking hawthorn, or salting impolitely raises it. At 3, fae gain a free ``Thorn Blight'' SB; reduce by copper rites or gifts. The gifts must be given with the right hand. The right hand is the hand that does not hold iron.
  \item \textbf{Hunt Phases:} \emph{Sounding} (fear checks; pass to gain +1 die to pace), \emph{Coursing} (paths shorten for counted strides), \emph{Closing} (only truth-payment or clemency halts the circle). The truth-payment is always a secret. The secret is always about a door you left open.
  \item \textbf{Lair Actions (pick two):} \emph{Fold the Conversation} (re-target a statement); \emph{Shift the Count} (odd $\leftrightarrow$ even penalties); \emph{Petal Cut} (apply 1 Harm unless copper shown); \emph{Wrong Tide} (swap entry/exit costs at a gate). The wrong tide is always the ninth tide. At the ninth tide, the gate smiles.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Aelinnel, power flows through courtesy, oath, and the hidden geometry of the land. A Stone Prince's authority rests on indexed fae treaties; a Green Knight's on the trials he has witnessed. There is no single crown, only the shared understanding that a broken promise cuts deeper than a sword – and that a promise kept is a door that opens both ways.

\textbf{For the GM:}  
Power in Aelinnel society is measured in hazel tokens, oath-beads, and the precision of a courtesy. Patronage often takes the form of guest-boughs, tide-path keys, or the Green Knight's escort – each tied to a specific court, grove, or obligation. To emphasise the cultural mechanics:
\begin{itemize}
\item Use \textbf{Two-Voice Courtesy} to reward players who pay attention to the unspoken. The unspoken is where the fae live.
\item Enforce \textbf{Copper over Iron} – carrying iron is not forbidden, but it is a statement. The statement is always heard. The hearing is always remembered.
\item Let \textbf{The Unsaid Clause} be a source of future complications; the fae love what is not said. They will fill the silence with their own words – and their words have weight.
\item Track \textbf{Obligation to the Green Gate} as a persistent clock. Each crossing without proper payment ticks it toward a visit from the Huntsman. The Huntsman is always polite. The Huntsman is always hungry.
\end{itemize}
In Aelinnel society, the stone remembers every step – and the Green Gate forgets no promise. The ninth step is not a step. The ninth step is a door. The door is always open. You just have to remember how to knock.
\end{tcolorbox}

% Index entries
\index{Aelinnel|see{Stone, Bough, and Bright Things}}
\index{Stone, Bough, and Bright Things generator}
\index{Gnomes!Aelinnel}
\index{Faekin!Aelinnel}
\index{Places!Aelinnel}
\index{People!Aelinnel}
\index{Complications!Aelinnel}
\index{Rewards!Aelinnel}
\index{Green Gate}
\index{Thorn Court}
\index{Lady of Thorns}
\index{Green Knight}
\index{Stone Prince of Aelinnel}
\index{Huntsman of the Moonlit Ride}
\index{Elf-causey}
\index{Moonwell}
\index{Dolmen}
\index{Glamour fog}
\index{Stone-wight}
\index{Wild Hunt}
\index{Oath-bead}
\index{Green Market}
\index{Two-Voice Courtesy}
\index{Copper over iron}
\index{Thorn-Courteous alignment}
\index{Market-Square alignment}
\index{Forester-Trusted alignment}
\index{The ninth step}
\index{The Hollow}