\section{Aelinnel Regional Generator}
\label{chap:aelinnel}

\textbf{Key Demographic: } Aelinnel (Gnomes)
\textbf{Captial: } Aevrossa

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{Count to eight. Stop. The ninth is for the Hollow.}

\begin{quote}
  \textbf{Bell-Warden of the Seventh Tide (Elite):}  
  ``The Hollow is not a place. It is a miscount. The ninth step that is never there, the ninth cup that is never poured, the ninth name that is never spoken. We do not fight the Hollow. We leave one empty chair, one unlit lamp, one uncounted stone. That is the price of sitting at the table.''
\end{quote}

\begin{quote}
  \textbf{Tide-Counter (Commoner):}  
  ``I asked an Aelinnel bell-tender how she kept the tide-spirits from flooding her village. She tapped her temple. \textit{I count them,} she said. \textit{One, two, three, four, five, six, seven, eight. Then I stop. They know that if I ever count nine, I am not counting them.''}
\end{quote}

\textbf{Genre / Mood:} Folk mathematics, coastal spirit-bargaining, horror of the miscount.

\textbf{Starting Location:} A floating shrine on the Inland Sea, where copper bells hang from every corner. A Bell-Warden rings a bell eight times, then stops: ``The tide-spirit is restless. It wants a ninth bell. I will not give it. You --- count the waves. Tell me how many break before the sun sets. If you miscount, the spirit will ask you for a story. Do not miscount.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
The Aelinnel of the East do not hide from the spirits --- they coexist. Small spirits inhabit every rock, tree, and wave. The Aelinnel have learned to negotiate with them using copper bells, precise counting, and occasional offerings of pickled plums. They are the engineers of shrines, the builders of floating bridges, and the keepers of the bell-lines that keep tide-spirits from straying. Some say Kuva, the sky, taught them the first counting song, and that the Hollow is simply the space where the count fails. Unlike their western kin, who lean toward the Gift of the Mind, the Eastern Aelinnel practice a folk mathematics that blends logic and superstition. They believe every spirit has a number-name --- a frequency that can be rung out of a bell, a sequence of steps that can be walked, a knot that can be tied. To offend a spirit is to miscount.

\vspace{2mm}
\textbf{What you notice first:} The sound of copper bells --- not ringing in sequence, but chiming in overlapping waves. The way every bridge has a missing ninth plank. The smell of pickled plums and salt.

\textbf{The one rule that kills newcomers:} Never count to nine aloud near a shrine. The spirits count with you. When you reach nine, they will ask: ``What comes after nine?'' The answer is always the Hollow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Count}

Power here is measured in bells rung, tides predicted, and the precision of a threshold. The Aelinnel have no kings or councils; their authority rests on the bell-wardens who maintain the shrines and the tide-counters who read the waves. A village that loses its bell-warden is a village that the tide-spirits will flood. A bridge without its ninth plank missing is a bridge that the Hollow will cross. The Aelinnel do not worship the spirits; they transact with them. An offering is not a prayer --- it is a payment. A bell is not a musical instrument --- it is a contract.

\subsection*{Prominent NPCs of the Aelinnel}
\label{sec:aelinnel-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Bell-Warden Tumble & Keeper of the Seventh Tide & Keep the bell-lines ringing and the Hollow at bay & He writes letters on copper wax; the wax is his only memory. \\
\hline
Tide-Counter Mira & Reader of the ninth wave & Predict the flood before it comes & She counts on her fingers but has only eight. The ninth is a phantom. \\
\hline
Shrine-Builder Orin & Engineer of floating shrines & Build a bridge that never needs a missing plank & He uses no nails; the wood is joined by counting songs. \\
\hline
Spirit-Negotiator Luma & Bargainer with tide-spirits & Trade pickled plums for safe passage & She has never lost a negotiation. The spirits fear her. \\
\hline
The Miscount (Ghost) & A bell-warden who counted to nine & Warn others of the Hollow's attention & It appears as a flickering shadow, holding one copper bell that never rings. \\
\hline
\end{longtable}

\subsection*{Names of the Aelinnel}
\label{sec:aelinnel-names}

Inspired by the inland sea cultures: short, melodic, often with doubled consonants. Surnames denote a tide, a shrine, or a counting song.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/Epithet} & \textbf{Bell-Name} \\
\hline
Tumble & Mira & Seventh Tide & \textit{the Nine-Counter} \\
\hline
Orin & Luma & Copper-kin & \textit{Bell-Born} \\
\hline
Kellen & Nori & Wave-reader & \textit{the Unrung} \\
\hline
Pip & Suki & Shrine-hand & \textit{Ninth Plank} \\
\hline
Rinn & Tali & Counting-song & \textit{the Still Bell} \\
\hline
\end{longtable}

\paragraph*{(Shrine/Bridge/Cove)} A floating shrine with copper bells; a bridge with a missing ninth plank; a spirit cove where offerings are left.

\section*{Spades --- Places}
\label{sec:aelinnel-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Floating Shrine (Copper Bells)} --- A wooden platform on pontoons, hung with bells.
  \hfill \textit{The bells ring when a spirit passes. The tone tells if the spirit is friendly.} \texttt{[SPIRIT]}
\item \textbf{Bridge of the Missing Plank} --- A wooden bridge with eight planks. The ninth is absent.
  \hfill \textit{The gap is exactly one foot wide. Step over it. Do not step in it.} \texttt{[THRESHOLD]}
\item \textbf{Spirit Cove (Offering Rocks)} --- A sheltered beach where offerings are left at low tide.
  \hfill \textit{The offerings are rice, pickled plums, and small copper coins. The spirits take them at night.} \texttt{[SPIRIT]}
\item \textbf{Bell-Line Tower} --- A stone tower with a rope of bells stretching across the channel.
  \hfill \textit{The rope is made of braided hair and seaweed. The bells are tuned to the tide's frequency.} \texttt{[BELL]}
\item \textbf{Counting School} --- A wooden hut where children learn the eight-number system.
  \hfill \textit{The teacher has only eight fingers. She was born that way. The children count on theirs.} \texttt{[SOCIAL]}
\item \textbf{Tide-Watcher's Post} --- A high platform on stilts, overlooking the sea.
  \hfill \textit{The watcher marks each wave with a pebble. Nine pebbles mean the tide is turning.} \texttt{[WEATHER]}
\item \textbf{Shrine of the Uncounted} --- A small shrine with no bells, no offerings, no threshold.
  \hfill \textit{It is for the spirits that have no number. They are the most dangerous.} \texttt{[DEATH]}
\item \textbf{Plum Orchard (Pickled)} --- A grove of plum trees, the fruit fermented in salt brine.
  \hfill \textit{The pickling barrels are sealed with copper lids. The lids have eight holes.} \texttt{[TRADE]}
\item \textbf{Copper Foundry} --- A forge where bells are cast and tuned.
  \hfill \textit{The foundry is silent. The workers use hand signals. Sound would disturb the spirits.} \texttt{[CRAFT]}
\item[J] \textbf{The Ninth Bell (Forgotten)} --- A bell that was cast but never hung. It sits in a field, overgrown.
  \hfill \textit{It rings once a year, on the night of the ninth tide. No one knows why.} \texttt{[MYSTERY]}
\item[Q] \textbf{Hollow's Gate (Submerged)} --- A stone arch underwater, visible only at the lowest tide.
  \hfill \textit{The arch is covered in counting marks. Eight marks, then a space. The space is the gate.} \texttt{[DEATH]}
\item[K] \textbf{The Counting Stone} --- A flat rock carved with eight grooves. Water flows through them.
  \hfill \textit{The ninth groove is missing. The water falls into a hole. The hole is bottomless.} \texttt{[MAGIC]}
\item[A] \textbf{The First Bell (Myth)} --- A bell that rang once and shattered. Its shards are scattered across the Inland Sea.
  \hfill \textit{Find all nine shards, and you can ring the bell that ended the first war with the Hollow.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Warden/Counter/Builder)} Bell-Warden Tumble with a copper bell; Tide-Counter Mira with eight fingers; Shrine-Builder Orin with a counting song.

\section*{Hearts --- People \& Factions}
\label{sec:aelinnel-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Bell-Warden} --- Old, gray-haired, a copper bell on a leather thong. He never speaks above a whisper.
  \hfill \textit{``The spirits hear everything,'' he says. ``A whisper is a negotiation. A shout is a declaration of war.''} \texttt{[SOCIAL]}
\item \textbf{Tide-Counter} --- A young woman with eight fingers. She counts waves by tapping her knee.
  \hfill \textit{``The ninth wave is the spirit's breath,'' she says. ``Do not count it. It counts you.''} \texttt{[WEATHER]}
\item \textbf{Shrine-Builder} --- A man with calloused hands and a wooden mallet. He never uses nails.
  \hfill \textit{``Nails are for coffins,'' he says. ``Joinery is for shrines. The spirits respect wood grain.''} \texttt{[CRAFT]}
\item \textbf{Spirit-Negotiator} --- A woman with a basket of pickled plums. She speaks to the air.
  \hfill \textit{``I offer eight plums,'' she says. ``The ninth is for later. Do you accept?'' The air ripples.} \texttt{[SOCIAL]} \texttt{[SPIRIT]}
\item \textbf{Copper-Smith (Silent)} --- A dwarf-like figure who works in complete silence. He uses hand signs.
  \hfill \textit{His hands are scarred from molten copper. He has not spoken in forty years.} \texttt{[CRAFT]}
\item \textbf{Tide-Spirit (Visible)} --- A shimmer of light over the water, shaped like a woman.
  \hfill \textit{It asks: ``How many waves have you counted today?'' Answer with eight, or it will flood.} \texttt{[SPIRIT]} \texttt{[FEAR]}
\item \textbf{Counting Teacher} --- An old woman with a slate and chalk. She teaches children to stop at eight.
  \hfill \textit{``What comes after eight?'' she asks. The children chorus: ``Nothing. We stop.''} \texttt{[SOCIAL]}
\item \textbf{Hollow-Watcher} --- A boy who sits at the edge of the bridge at dusk, watching the missing plank.
  \hfill \textit{He can see the Hollow when it crosses. He whistles to warn the village.} \texttt{[DEATH]}
\item \textbf{Offering-Keeper} --- A woman who tends the spirit cove. She sweeps the beach at dawn.
  \hfill \textit{``The offerings are gone,'' she says. ``The spirits were satisfied. We have one more day.''} \texttt{[SPIRIT]}
\item[J] \textbf{The Miscount's Ghost} --- A flickering shadow with a copper bell that never rings.
  \hfill \textit{It follows travelers who count to nine. It whispers: ``Nine. Nine. What comes after nine?''} \texttt{[DEATH]} \texttt{[FEAR]}
\item[Q] \textbf{First Bell's Shard (Keeper)} --- A woman who holds one of the nine shards. She guards it with her life.
  \hfill \textit{The shard is warm. It hums at low tide. It wants to be reunited.} \texttt{[MAGIC]}
\item[K] \textbf{The Still Bell (Legend)} --- A bell that hangs in a submerged shrine. It rings only when the Hollow walks.
  \hfill \textit{No one has heard it ring in a century. The Hollow has been quiet. Too quiet.} \texttt{[LEGEND]}
\item[A] \textbf{Kuva's Echo (Sky Spirit)} --- A voice from the bell-line tower, speaking in counting songs.
  \hfill \textit{``You have counted eight,'' it says. ``That is enough. The ninth is mine.''} \texttt{[PATRON]}
\end{enumerate}

\paragraph*{(Squall/Pirate/Whirlpool)} A sudden squall --- all navigation rolls at disadvantage; a pirate attack --- pay cargo or fight; a spirit-whirlpool --- lose 1 Supply to escape.

\section*{Clubs --- Complications/Threats}
\label{sec:aelinnel-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Sudden Squall} --- A storm rises from calm waters. Navigation DV +2. \texttt{[WEATHER]}
  \hfill \textit{The squall was summoned by an offended tide-spirit. Someone miscounted.}
\item \textbf{Pirate Attack (Kalingan)} --- Fast dhows appear from behind a reef. They demand cargo. \texttt{[COMBAT]}
  \hfill \textit{The pirates are not human. They are spirit-touched. They count in nines.}
\item \textbf{Spirit-Whirlpool} --- A whirlpool opens beneath the boat. Lose 1 Supply to escape. \texttt{[DEATH]} \texttt{[TRAVEL]}
  \hfill \textit{The whirlpool is a spirit's mouth. It is hungry. Feed it a memory.}
\item \textbf{Bell-Line Breaks} --- A rope of bells snaps. The tide-spirits are no longer bound. \texttt{[SPIRIT]} \texttt{[TOLL]}
  \hfill \textit{The tide rises. The village will flood in nine hours. Repair the line.}
\item \textbf{Miscount (Offended Spirit)} --- A traveler counted to nine near a shrine. The spirit demands compensation. \texttt{[SOCIAL]} \texttt{[TOLL]}
  \hfill \textit{The compensation is a story that ends with the number eight. The spirit will check.}
\item \textbf{Missing Offering} --- The spirits took the offering, but they left a mark. The mark is a warning. \texttt{[OMEN]}
  \hfill \textit{The warning says: ``The Hollow is coming. Ring the ninth bell.'' But the ninth bell does not exist.}
\item \textbf{Copper Foundry Fire} --- The silent forge catches fire. The bells crack. The tones are wrong. \texttt{[DISASTER]}
  \hfill \textit{The fire was set by a rival bell-warden. He wants to replace the old bells with his own.}
\item \textbf{Bridge Plank Replaced (By Mistake)} --- Someone nails a ninth plank onto the bridge. The Hollow crosses. \texttt{[DEATH]}
  \hfill \textit{The Hollow asks: ``Who built this bridge?'' The builder is the party's ally.}
\item \textbf{Counting School Poisoned} --- The teacher's chalk is laced with a memory-erasing dust. \texttt{[MAGIC]} \texttt{[TOLL]}
  \hfill \textit{The children forget to stop at eight. They count to nine. The Hollow comes for them.}
\item[J] \textbf{The Ninth Bell Rings (Once)} --- The forgotten bell in the field rings. Everyone hears it. \texttt{[OMEN]} \texttt{[DEATH]}
  \hfill \textit{The Hollow's gate opens. Spirits flood the Inland Sea. The Aelinnel are terrified.}
\item[Q] \textbf{Tide-Spirit Demands a Name} --- A spirit rises from the water. It asks: ``What is the ninth number?'' \texttt{[RIDDLE]} \texttt{[DEATH]}
  \hfill \textit{There is no ninth number. The correct answer is silence. The wrong answer is death.}
\item[K] \textbf{First Bell Shard Stolen} --- Someone steals one of the nine shards. The Hollow stirs. \texttt{[THEFT]} \texttt{[TOLL]}
  \hfill \textit{The thief is a foreign merchant. He wants to sell the shard to the Veiled Sisters.}
\item[A] \textbf{Kuva's Silence} --- The counting song stops. The sky is empty. No birds, no wind, no waves. \texttt{[APOCALYPSE]}
  \hfill \textit{Kuva has turned away. The Hollow is free. The Aelinnel are counting their last breaths.}
\end{enumerate}

\paragraph*{(Charter/Bead/Stone)} Bell-charter (right to ring a named shrine); tide-bead (safe passage on the Inland Sea for one crossing); counting stone (one reroll on a navigation test in fog).

\section*{Diamonds --- Treasures \& Discoveries}
\label{sec:aelinnel-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Bell-Charter (Copper Token)} --- A small copper disc. Right to ring any named shrine bell once. \texttt{[SOCIAL]}
  \hfill \textit{The token is stamped with eight dots. The ninth dot is missing.}
\item \textbf{Tide-Bead (Glass)} --- A blue glass bead. Safe passage on the Inland Sea for one crossing. \texttt{[TRAVEL]}
  \hfill \textit{The bead was blessed by a tide-spirit. It glows when a storm is coming.}
\item \textbf{Counting Stone (Smooth Pebble)} --- A pebble with eight grooves. Reroll a failed navigation test in fog. \texttt{[TRAVEL]}
  \hfill \textit{The stone was worn by waves for a century. It knows the tide's schedule.}
\item \textbf{Offering Box Key (Copper)} --- A small key. Opens any Aelinnel offering box. \texttt{[STEALTH]}
  \hfill \textit{The key is warm. It glows when an offering has been accepted.}
\item \textbf{Shrine-Builder's Mallet} --- A wooden mallet with no nails. Build a temporary shrine anywhere. \texttt{[CRAFT]}
  \hfill \textit{The shrine lasts one night. The spirits will respect it.}
\item \textbf{Spirit-Negotiator's Plum (Pickled)} --- A pickled plum in a brine jar. Offer it to any tide-spirit. \texttt{[SOCIAL]}
  \hfill \textit{The spirit will grant one favor. The favor is never large. It is always fair.}
\item \textbf{Bell-Line Rope (Hair and Seaweed)} --- A length of braided rope. Use it to bind a tide-spirit for one hour. \texttt{[MAGIC]}
  \hfill \textit{The rope smells of salt. The spirit will remember the binding.}
\item \textbf{Tide-Counter's Pebble Pouch} --- A leather pouch of marked pebbles. Predict the tide for the next day. \texttt{[WEATHER]}
  \hfill \textit{Count the pebbles. Eight mean high tide. Seven mean low. Nine mean run.}
\item \textbf{Copper Bell (Untuned)} --- A small bell that has not been tuned. Ring it to confuse a spirit. \texttt{[COMBAT]}
  \hfill \textit{The spirit will hesitate for one exchange. The bell will crack after use.}
\item[J] \textbf{Forgotten Bell's Tone (Recording)} --- A clay cylinder that captured the ninth bell's ring. \texttt{[MAGIC]}
  \hfill \textit{Play it to open the Hollow's gate. Use once. The Hollow will be angry.}
\item[Q] \textbf{First Bell Shard (Green Glass)} --- A shard of the mythic first bell. Touch it to any spirit to command obedience for one minute. \texttt{[MAGIC]}
  \hfill \textit{The shard is warm. It hums. The spirit will obey --- but it will remember.}
\item[K] \textbf{Kuva's Counting Song (Flute)} --- A bone flute that plays the eight-note counting song. \texttt{[SOCIAL]}
  \hfill \textit{Play it to calm any spirit. The spirit will listen. It will not attack.}
\item[A] \textbf{The Hollow's Silence (Bell)} --- A copper bell that rings only for the Hollow. Ring it to banish an Oath-Keeper. \texttt{[DEATH]}
  \hfill \textit{The bell will ring once. Then it will shatter. The Hollow will not forget.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:aelinnel-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = treasure. Highest rank sets the main Timer (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Treasures (Diamonds) are codified outcomes (charters, beads, stones) that change position rather than call for a roll.
\item If any A appears, echo \textbf{bell-omens} --- a bell that rings without being struck, a missing plank that appears, a count that reaches nine and stops.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:aelinnel-mechanics}

\begin{itemize}
\item \textbf{Bell-Tending:} Once per scene, when near a shrine or threshold, you may ring a copper bell. The next environmental hazard (tide, fog, landslide) is delayed by 1 segment of the relevant timer. The bell's tone also pacifies minor spirits (Cap 1) for the scene.
\item \textbf{Fae Courtesy:} An Aelinnel never says ``no'' to a spirit. They say ``that road is closed for repairs'' or ``the season is not right.'' This phrasing cancels the first Story Beat generated by a spirit encounter. If forced to refuse outright, they mark 1 Fatigue (the spirit's disappointment is heavy).
\item \textbf{Kuva's Count:} Once per journey, a character may roll Wits + Lore (DV 3) to recall a forgotten sky-omen; on success, the next travel leg ignores the first Club card drawn. On a miss, the omen is true, but the GM draws an additional Club for that leg.
\item \textbf{The Ninth Step:} When you cross a threshold that has a missing ninth step (a stair, a bridge, a door), you may count the absence. Gain +1 die to resist the Hollow's attention for the scene, but mark 1 Obligation (the Hollow now knows your name).
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Unrung} --- The Witness. \textit{A bell that never rings, yet remembers every promise.}
    \item \textbf{The Counter} --- Mykkiel. \textit{The divine arithmetician; a miscount is a sin.}
    \item \textbf{The Ninth Wave} --- Khemesh. \textit{The pressure beneath the tide; the hunger that waits after eight.}
    \item \textbf{Kuva's Echo} --- The Traveler. \textit{The sky-road and the counting song that guides lost ships.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Bitter Root:} Daughters of the Cord sometimes serve as bell-wardens, using the Hollow's attention to collect on promises. A renegade Sister hides in a floating shrine, trying to untie her cord from a tide-spirit's bargain.
  \item \textbf{Malachai, the Chained Angel:} A silver coin that never spends appears in offering boxes. Take it, and the spirits will accept any gift --- but the coin will reappear in the mouth of a drowned sailor.
  \item \textbf{The Gray Wanderer's Grimoire:} Aelinnel psions are called ``Echo-Readers.'' They hear the count of waves in their dreams and can predict a miscount three tides before it happens.
  \item \textbf{The Threshold Compendium:} The Ninth Bell is a Class III Procession -- the sound of the bell walks the Inland Sea once a year, and those who hear it must answer a riddle or join the march.
  \item \textbf{The Ninth:} In any Aelinnel scene, the ninth bell rung in an hour opens a threshold. The wise stop at eight. Those who ring the ninth see a gray figure standing on the water --- waiting.
\end{itemize}

\subsection*{Regional Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Bell:} Crumble a crust at the base of a shrine bell. Ignore the first \emph{Bell-Line Breaks} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Rope:} Tie a single knot in a bell-line rope. Cancel \emph{Miscount} or take --1 die on your next spirit negotiation (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Wave:} Whisper a dead tide-counter's name into the sea. Gain +1 Effect when counting waves this scene, but mark \emph{Obligation} to Kuva. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:aelinnel-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{bell-wardens or tide-spirits}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{shrine builders or offering-keepers}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{spirits or the Hollow}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local Bell-Warden or tide-counter} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{fog or rising water}.
\item \textbf{Foe Appears:} An \textbf{offended tide-spirit or Hollow agent} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{spirits, bell-wardens, or Hollow manifestations} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes --- \textbf{bell-line breaks/miscount called/ninth bell rings}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Bell-Warden Tumble, Kuva's Echo, or the Miscount's Ghost} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically --- \textbf{shrine floods/bell-line snaps/ninth plank appears}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice --- \textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Aelinnel (Bell Law):} Bells ring in harmony, then stop; a missing plank glows; the count of waves becomes audible.
\item \textbf{Aelinnel (Spirit Law):} A tide-spirit laughs, an offering steams, a copper bell tarnishes instantly.
\item \textbf{Aelinnel (Hollow Law):} The ninth step appears, then vanishes; a whisper says ``count again''; a shadow crosses the bridge.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Navigation):} On a hit, shift one band. On a strong hit, also impose \emph{Fog Bank}.
\item \textbf{Cut the Bell (Presence + Lore):} On a hit, freeze range for one exchange; if a bell-charter is in your possession, +1 die.
\item \textbf{Weather the Squall (Resolve + Survival):} On a hit in \emph{Sudden Squall}, you pick who keeps their footing; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Bell \& Count)}
\begin{itemize}
\item \textbf{The Miscount-King (Nine Errors)} --- \emph{Phases}: Bell-Lines Severed $\rightarrow$ Orchestrated Miscount $\rightarrow$ Shrine Coup. \emph{Cracks}: re-tie a bell-line, expose a false count, survive a night at the forgotten bell.
\item \textbf{The Drowned Counter (Spirit of the Ninth Tide)} --- \emph{Phases}: Waves Miscounted $\rightarrow$ Procession of Empty Shrines $\rightarrow$ Bell-Baptism of a Warden. \emph{Cracks}: relic tide-bead, survivor testimony, the ninth bell refusing to ring.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power},breakable]
Among the Aelinnel, power flows through bells, counts, and the precision of a threshold. A Bell-Warden's authority rests on the number of shrines he tends; a Tide-Counter's on the accuracy of her predictions. There is no government, only the shared understanding that a miscount kills everyone.

\textbf{For the GM:}  
Power in Aelinnel society is measured in bells rung, offerings given, and miscounts avoided. Patronage often takes the form of bell-charters, tide-beads, or counting stones --- each tied to a specific shrine, tide, or spirit. To emphasize the cultural mechanics:
\begin{itemize}
\item Use \textbf{Bell-Tending} as a hazard mitigation tool. Let players delay environmental timers by ringing bells.
\item Enforce \textbf{Fae Courtesy} --- never say ``no'' to a spirit. Creative phrasing is rewarded.
\item Let \textbf{Kuva's Count} be a journey-saving ability, but with a risk of additional complications.
\item Track \textbf{The Ninth Step} as a source of Hollow Obligation. Each crossing of a threshold is a negotiation.
\end{itemize}
In Aelinnel society, the count forgets nothing --- and the ninth bell is never rung.
\end{tcolorbox}

% Index entries
\index{Aelinnel|see{Bell-Tenders}}
\index{Theme generator}
\index{Bell-Tenders generator}
\index{Places!Aelinnel}
\index{People!Aelinnel}
\index{Complications!Aelinnel}
\index{Rewards!Aelinnel}
\index{Bell-tending}
\index{Counting songs}
\index{Kuva's count}
\index{Ninth step}
\index{Fae courtesy}
\index{Tide-spirits}
\index{Bell-lines}
\index{Forgotten bell}
\index{Local Patrons!Aelinnel}
