\section{Ashaan --- ``The Veiled Throne''}
\label{chap:ashaan}

\begin{tcolorbox}[colback=red!5!white,colframe=red!60!black,title={Content Warning}]
Ashaan is a setting of systemic cruelty: chattel slavery, state-sanctioned violence, child sacrifice (secret cults), and the erasure of identity. The Veiled Sisters rule through terror, the Black Hand erases dissent, and the Breath-less resistance struggles in shadows. These themes are presented as horror and injustice, never glamourised. Before playing, use Lines \& Veils to ensure everyone's comfort. If you need to adjust or omit elements, the GM has final authority. Your safety matters more than canon.
\end{tcolorbox}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The Veil is not a choice. It is survival. Speak in thirds — two options and a hidden third. The ninth word is always a knife.}

\begin{quote}
  \textbf{Veiled Sister (Elite):}  
  ``The Veil is not a choice. It is survival. You will show your eyes and nothing more. Speak in thirds — offer two options and a hidden third. The ninth word is always a knife.''
\end{quote}

\begin{quote}
  \textbf{Breath-less Agent (Commoner):}  
  ``They say Ikasha sleeps. I say she is patient. The Maeri have ruled for generations; they believe the Veil is eternal. But the Veil is silk. Silk frays. And I am learning to bite through — one whisper at a time.''
\end{quote}

\noindent\textbf{Genre / Mood:} Grimdark spirit-horror, bureaucratic terror, whispered resistance, creeping divine punishment.

\noindent\textbf{Starting Location:} A root cellar beneath a spice merchant's shop, lit by a smuggled candle. A Breath-less agent presses a folded paper into your hand: ``My brother argued with an Esoiti about a misweighed sack of cardamom. His name was erased from the rolls by morning. Help me find him — or help me remember his name.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Once, the Ashaani Empire stretched across the eastern Amaranthine rim — a river-fed dominion of sorcerer-kings, the 999 Immortals, and the great twin-river metropolis of Khasurah. Three centuries ago, an invasion of Oshiira shattered the empire. The Immortals were annihilated; plague, blockade, and northern mercenaries finished the ruin. Today, what remains is a rump city-state called Ashaan (or As'Ashaan), a fortified river-island north of the Khesai delta. The old lands — the \textbf{Green Ribbon} floodplains, the \textbf{Great Desert} where glass hums with ancient curses — are a patchwork of loyal towns, restless hill peoples, and the ruins of swallowed cities like Khasurah, which still breathes under dunes and salt crusts.

But the fall of Ashaan was not merely conquest. It was \emph{punishment}. The priests whisper that the 999 Immortals committed a sin so profound that the gods turned their faces away — and the sand remembers what the living have forgotten. The Veil, the magical membrane that separates the living from the dead, is not a barrier. It is a \emph{wound} that never healed. And the wound is spreading.

\textbf{Neighbors and Rivals:}
\begin{itemize}
  \item \textit{West – Oshiira:} The grain empire that broke Ashaan's back. The two nations maintain a cold peace, punctuated by Oshiiran grain barges that pay ruinous tariffs to pass the Brass Gate. Ashaani spies in Mbaro watch for any sign that the Widow might seek to finish what her ancestors started.
  \item \textit{South – Taharka:} The highland kingdom of the Monsoon Crown. Taharkan passes are a graveyard for Ashaani armies; the Veiled Sisters have never broken those mountains. Now Ashaani agents slip through disguised as spice merchants, seeking to weaken Taharka from within.
  \item \textit{East – Galanina (Sidhi ancestral homeland) and the Pereshi Highlands:} Galanina, the old eastern capital, lies under Ashaani control — but its Catacombs of the Unburied hold the bound spirits of Sidhi slaves, and the Breath-less dream of freeing them. Beyond, the Pereshi fire temples guard the secret of the Forgotten Prince, an Ashaani conqueror whose true name was severed. The Maeri want that name; the Pereshi will never give it.
\end{itemize}

The state is ruled by three masked beings known as the \textbf{Maeri} — the Veiled Sister (coins and winter), the Helmed Sister (steel and sickness), and the Masked Sister (red and black). No one knows what they are: demigods, archfey, lich-queens, or the last masks of the First Fey. They speak through the \textbf{Esoti}, a collegium of wizards who maintain the wards, govern the bureaucracy, and interpret the Maeri's will. In secret, certain Esoti circles conduct child sacrifice to Khemesh, a sea-demon cult imported in the empire's maritime days.

The \textbf{Black Hand} is the Maeri's instrument: a fanatical network of bards and warlocks divided into three Aspects — Veiled (accountants, heralds, ruin by paper), Helmed (assassins and enforcers, ruin by blade), and Masked (courtesans, impresarios, ruin by desire). The Hand removes over-mighty nobles, disappears rival magi, and unpersons threats. Within it, ambition is a short road to an unmarked grave. Unlike the Sisters, the Hand is \emph{human} — broken, desperate, or fanatical, but still capable of fear, doubt, and the occasional flicker of mercy.

Slavery is legal, widely practiced, and increasingly focused on wandered Sidhi lineages broken by the empire's fall. Enslaved Sidhi bear blue marks on the forearm; illegal to remove without a manumission writ, but earth-wise midwives know ways to fade it. In the catacombs beneath Galanina — the old eastern capital — the dead are mapped like a city, and the spirits of the unburied are bound into brass cages.

Yet the Veil is cracking. In Galanina, devotion to the Angel Mykkiel (a solar of unbending Law and Justice) spreads. His zealots are incorruptible — an acid slowly eating the old order. And beneath the capital, the goddess Ikasha (She Who Sleeps) stirs in her sealed vault. Her followers, the \textbf{Breath-less}, meet in root cellars, learn to read records without leaving marks, and whisper that a name restored can weaken the Veil.

\vspace{2mm}
\textbf{What you notice first:} The smell of incense and rust. The way every doorway has a veil of brass chains. The silence when a Veiled Sister passes — even the spirits hold their breath. The brass tags on the dead — one for each Sister. And the green glass that glints in every shadow, fragments of a desert that was once a garden.

\textbf{The one rule that kills newcomers:} Never name the dead. The Sisters bind spirits to names. Speak a dead name aloud, and you may find yourself bound alongside them. Also, never address a coin-veiled speaker at court unless you are ready to accept their price.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of Damnation}

Power here is measured in spirits bound, names erased, and the silence of witnesses. The Maeri rule through terror; the Esoti hold the records; the Black Hand hunts without uniform. The Breath-less whisper in cellars, learning to read records without leaving marks. Ashaan is not a place of open rebellion. It is a place of whispered warnings, vanished neighbors, and scrolls that rewrite themselves.

But beneath all of this runs the \emph{creeping}. The Red Desert expands by inches each year, its green glass humming with a forgotten curse. The spirits bound in brass cages grow restless; their whispers are prayers to gods who no longer answer. The Veil, once a clean wound, has begun to fester. No one remembers what sin caused it. Perhaps no one ever knew. But everyone feels the weight — a pressure behind the eyes, a dryness in the throat, a certainty that something is \emph{coming}.

\subsection*{Prominent NPCs of Ashaan}
\label{sec:ashaan-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
The Veiled Sister & Maeri of coins and winter & Maintain the Veil and the tribute system & She has never removed her veil of gold coins. The coins are from dead empires. She does not blink. Her voice is the same as the wind over a mass grave. \\
\hline
The Helmed Sister & Maeri of steel and sickness & Enforce order through terror & Her bronze helm never opens. Those who see inside never speak again. Her shadow has teeth. She does not eat. \\
\hline
The Masked Sister & Maeri of red and black & Weave desire into control & She knows every secret whispered in Ashaan's brothels and art houses. Her smile is a contract. She has no reflection. \\
\hline
Magistrate Eso & Esoiti high magistrate & Control every record, every court, every gate & He rewrites his own memory each morning. His left hand is a stamp. His right hand is a key that fits no lock. \\
\hline
The Black Hand (Aspect Commander) & Faceless hunter & Erase anyone who whispers Ikasha's name & No one knows their faces. They could be your neighbor, your clerk, your spouse. They are never seen eating. \\
\hline
Tamar & Breath-less cell leader & Tear down the Veil from within & She has three names; each is for a different jurisdiction. Her shadow moves only when she is still. \\
\hline
Kaelen & Spirit-broken slave & A child taken for ``reevaluation,'' returned hollow-eyed & He no longer speaks. He only weeps when he hears a brass bell. His shadow is a brass cage. \\
\hline
The Whisperer & Forbidden mage in hiding & Learn to read records without leaving marks & He writes on water. The words vanish, but the Sisters cannot read them. His reflection is missing. \\
\hline
\end{longtable}

\subsection*{Names of Ashaan}
\label{sec:ashaan-names}

Inspired by the eastern river-core and the long shadow of old empires: sharp consonants, vowel-rich endings, with echoes of ancient Egyptian and Seljuk naming conventions. Surnames often denote a lost lineage, a trade, or a place swallowed by the sand.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Kaelen & Morai & al-Galanina & \textit{the Veiled} \\
\hline
Tarek & Iman & Esoiti-broken & \textit{the Breath-less} \\
\hline
Jasin & Nadira & Black-Hand-marked & \textit{the Unnamed} \\
\hline
Zahir & Tamar & Spirit-bound & \textit{the Ninth Witness} \\
\hline
Rashid & Layla & Ledger-erased & \textit{the Forgotten} \\
\hline
Dawud & Amina & Cage-born & \textit{the Hollow} \\
\hline
Seti & Neferet & Khasurah-lost & \textit{the Glass-Voiced} \\
\hline
\end{longtable}

\paragraph*{(Gate/Sanctum/Pan)} Salt pan with a broken spirit-cage; black marble temple with steps stained red; the Brass Gate — a fortified strait with spirit-buoys.

\section*{\spadecard{A} Places}
\label{sec:ashaan-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Salt Pan with Broken Spirit-Cage} --- A white flat where the air hums. \texttt{[DEATH]} \texttt{[SPIRIT]}
  \hfill \textit{The cage is empty, but it still rattles at night. Something got out. The salt tastes of iron.}
\item \textbf{Black Marble Temple (Maeri)} --- Steps stained red; the blood never washes away. \texttt{[DEATH]} \texttt{[HORROR]}
  \hfill \textit{The temple has no windows. The Maeri do not need light. The steps are warm, as if a great beast sleeps beneath.}
\item \textbf{The Brass Gate} --- A fortified strait with spirit-buoys that scream when crossed. \texttt{[TRAVEL]} \texttt{[SPIRIT]}
  \hfill \textit{Each buoy contains a bound spirit. They sing the names of those who pass. The ninth buoy is silent.}
\item \textbf{Veiled Sister's Sanctum} --- Entrance hidden by a mirage that shows your greatest fear. \texttt{[MAGIC]} \texttt{[FEAR]}
  \hfill \textit{The mirage is not an illusion. It is a memory the Sister has stolen. The memory is always from a child.}
\item \textbf{Galanina Slave Market} --- A circular pit with cages of brass and bone. \texttt{[DEATH]} \texttt{[WEIGHT]}
  \hfill \textit{The auctioneer wears a mask. The mask is made from the face of a debt-seller. The face still moves.}
\item \textbf{Esoiti Record Hall} --- A long chamber with shelves of silent scrolls. \texttt{[LAW]} \texttt{[WEIGHT]}
  \hfill \textit{The scrolls write themselves. The ink is the blood of erased names. The ninth shelf is empty — reserved for the name that must never be recorded.}
\item \textbf{Black Hand Garrison} --- A nondescript building with no windows and a single door. \texttt{[DANGER]}
  \hfill \textit{The door has no handle. It opens inward only when a name is whispered. The whisper is never forgotten.}
\item \textbf{Breath-less Cellar} --- A root cellar with a single candle and a hole in the floor. \texttt{[SECRET]}
  \hfill \textit{The hole leads to a tunnel that leads to a tomb. The tomb is a safe house. The candle never burns lower.}
\item \textbf{Red Desert's Edge} --- Where the fertile land turns to dust and glass. \texttt{[WEATHER]} \texttt{[CURSE]}
  \hfill \textit{The glass is green. It hums with the memory of the prince's curse. The sound is the same as a brass cage rattling.}
\item[J] \textbf{The Ninth Name Shrine} --- A ruin where a forbidden name is carved into stone. \texttt{[DEATH]} \texttt{[SECRET]}
  \hfill \textit{The stone weeps water. The water tastes of iron and forgotten vows. The name is worn smooth — no one can read it, but everyone fears it.}
\item[Q] \textbf{Spirit-Binding Foundry} --- A forge where brass cages are made. \texttt{[CRAFT]} \texttt{[SPIRIT]}
  \hfill \textit{The forge is heated by a bound fire-spirit. It screams when the bellows pump. The screams are prayers in a dead language.}
\item[K] \textbf{The Hollow Vessel Nursery} --- An orphanage where children are prepared for possession. \texttt{[HORROR]} \texttt{[DEATH]}
  \hfill \textit{The children have no shadows. Their shadows have been sold. The ninth cot is always empty — for the child who will be the Hollow's vessel.}
\item[A] \textbf{Ikasha's Sleeping Chamber} --- A sealed vault beneath Galanina. \texttt{[MYSTERY]} \texttt{[PATRON]}
  \hfill \textit{Inside, a stone sarcophagus. The stone is warm. She Who Sleeps dreams of freedom. Her dreams leak through the cracks — green glass grows where they touch stone.}
\end{enumerate}

\paragraph*{(Hunter/Agent/Magician)} Slave-hunter (Black Hand) with a branding iron; Breath-less agent (freed slave, now spy); an Ashaani magician who offers dangerous knowledge.

\section*{\heartcard{A} People \& Factions}
\label{sec:ashaan-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Terrified Baker} --- Kneads dough with shaking hands. He knows which Black Hand guards take bribes and which take only silence. He will sell you bread, but he will not meet your eyes. If the Hand asks, he did not see you. He will weep afterward. \texttt{[SURVIVAL]} \texttt{[FEAR]}
  \hfill \textit{His son was taken for ``reeducation'' last year. He does not know if the boy is alive. He is afraid to ask.}
\item \textbf{Ledger Clerk (Esoiti)} --- Ink-stained fingers, hollow eyes. Copies erasure orders all day. He knows whose names are on next week's list. He will trade that knowledge for a single assurance that his own name isn't on it — but he cannot be sure you can deliver. He is not brave. He is merely desperate. \texttt{[LAW]} \texttt{[WEIGHT]}
  \hfill \textit{He has a wife and a daughter. The daughter's name is written on a scrap of paper hidden in his shoe. That is his only insurance.}
\item \textbf{Black Hand Recruit (Fresh)} --- Still has a family. Still flinches when the brand touches his arm. He joined because the mines are a death sentence. He does not know what he will be asked to do next week. He is terrified. He will let you pass if you promise to tell his mother he didn't suffer. \texttt{[COMBAT]} \texttt{[TRAGEDY]}
  \hfill \textit{He has never killed anyone. He does not know if he can.}
\item \textbf{Mother at the Well} --- Gave her youngest child's name to the Sisters. Not because she is evil. Because the alternative was her whole family being taken. She draws water at dawn when acolytes aren't watching. She will not speak of it. But if you ask about the \textit{ninth child}, she will weep. \texttt{[DEATH]} \texttt{[SORROW]}
  \hfill \textit{The child's name was Laila. She hums her daughter's lullaby to the rope. The well remembers.}
\item \textbf{Black Hand Veteran} --- Has served twenty years. Her hands do not shake. She has seen the Sisters' work up close. She knows there is no escape, only delay. She will not help you escape — but she might sell you a day's head start. The price is a story. The story must be true. \texttt{[DANGER]} \texttt{[WORLDWEARY]}
  \hfill \textit{She has forgotten the taste of honey. She would like to remember.}
\item \textbf{Esoti Junior Magistrate} --- Inherited her post from an uncle who disappeared. She knows the laws are lies. She enforces them anyway because the alternative is the Binding Foundry. She drinks too much and weeps in the archive. She has a key to a vault. She will lend it — for a story about what freedom tastes like. \texttt{[LAW]} \texttt{[SECRET]}
  \hfill \textit{The key is a brass finger bone. It fits a lock in the Ninth Ledger's vault. She does not know what it opens.}
\item \textbf{Slave (Household)} --- Not a rebel. Not a hero. Just a woman who has learned to be invisible. She knows the secret passages behind the kitchen. She will trade directions for a single piece of news from outside — any news, even bad. She wants to know if her village still exists. \texttt{[STEALTH]} \texttt{[SURVIVAL]}
  \hfill \textit{She wears a brass tag with a number. She has memorised the number of every slave in the household. She is waiting for the day the numbers no longer matter.}
\item \textbf{Black Hand Officer (Compromised)} --- He takes bribes. He looks the other way. He has a daughter in the Sisters' care as a ``ward.'' He knows she is a hostage. He will let you pass if you carry a letter to her — a letter that says nothing, but is written on paper that smells of home. He is not your ally. He is a man drowning. \texttt{[AUTHORITY]} \texttt{[TRAP]}
  \hfill \textit{The paper is his daughter's old schoolwork. The ink is his own blood. He does not know if she can still read.}
\item \textbf{Veiled Sister (Acolyte, not a full Sister)} --- She was given to the temple at nine. She does not remember her mother. She believes the Veil protects the empire from something worse. She has never seen a slave beaten — they hide that from novices. She is not evil. She is \emph{cultivated}. A crack in her certainty may take years to open. You do not have years. \texttt{[MYSTERY]} \texttt{[DANGER]}
  \hfill \textit{She wears a veil of hammered silver. Underneath, her face is smooth — no scars, no wrinkles. She has not aged since she took her vows.}
\item[J] \textbf{The Whisperer (Breath-less cell leader)} --- She has three names, each for a different jurisdiction. She moves through the catacombs like smoke. She will trust you only after you have broken a small law in her presence — not to prove your courage, but to prove you have something to lose. \texttt{[SECRET]} \texttt{[RESISTANCE]}
  \hfill \textit{She writes messages on water. The words vanish, but the Sisters cannot read them. Her reflection is missing. She sold it for a name.}
\item[Q] \textbf{The Black Hand's Shadow} --- A Hand agent who has stopped sleeping. He sees the Sisters' work in his dreams. He knows he cannot leave. He will not betray his oath — but he might fail to see you, if the price is a single night of dreamless rest. The price is a memory. He will take the happiest one. \texttt{[DEATH]} \texttt{[HORROR]}
  \hfill \textit{He carries a brass bell that rings once every hour. It is the only thing that keeps him awake. He has not heard it ring in three days.}
\item[K] \textbf{The Esoti Censor (Desperate)} --- A high-ranking magistrate who has been ordered to falsify a lineage. If she refuses, her family is erased. If she complies, she knows the truth will outlive her. She will give you the original document — for a promise that her grandson will be told, one day, why his grandmother's name was struck from the rolls. \texttt{[LAW]} \texttt{[TRAGEDY]}
  \hfill \textit{The document is written on human skin. The skin was her own, cut from her thigh when she took office. She has no feeling there anymore.}
\item[A] \textbf{The Veiled Sister (Full)} --- Malignant divine justice made flesh. She does not threaten. She does not bargain. She states what will happen. She is always correct. She has never been seen eating, sleeping, or blinking. She is not a woman. She is a function of the Veil. There is no complexity here. There is only the weight of a god's attention. \textit{(No debt. No price. Only survival.)} \texttt{[PATRON]} \texttt{[APOCALYPSE]}
  \hfill \textit{Her veil is made of gold coins from a dozen dead empires. They do not clink. They weep when she speaks.}
\end{enumerate}

\paragraph*{(Dust/Gate/Spirit)} Spirit-dust storm (Navigation DV +1, mark Fatigue); Brass Gate closed for ``inspection'' — delay 2 segments; a bound spirit breaks free and attacks.

\section*{\clubcard{A} Complications/Threats}
\label{sec:ashaan-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Spirit-Dust Storm} --- Red sand from the desert carries whispering spirits. \texttt{[WEATHER]} \texttt{[DEATH]}
  \hfill \textit{Navigation DV +1. Mark Fatigue. The voices ask for your name. Do not answer. They already know it.}
\item \textbf{Brass Gate Closed} --- The strait is sealed for ``inspection.'' Delay 2 segments. \texttt{[LAW]} \texttt{[DELAY]}
  \hfill \textit{The ``inspection'' is a hunt for a Breath-less agent. The agent is in your party. The gate knows. The gate is patient.}
\item \textbf{Bound Spirit Breaks Free} --- A brass cage cracks. The spirit attacks. \texttt{[DEATH]} \texttt{[COMBAT]}
  \hfill \textit{The spirit was once a child. It weeps as it fights. Killing it feels like murder. Freeing it feels like surrender.}
\item \textbf{The Veil Fails} --- For a moment, the Maeri's magic falters. Spirit-visions haunt all. \texttt{[MAGIC]} \texttt{[FEAR]}
  \hfill \textit{You see the face of someone you wronged. They are asking for an apology. You cannot remember what you did. That is the punishment.}
\item \textbf{Black Hand Sweep} --- A sudden search of the district. Everyone's papers are checked. \texttt{[LAW]} \texttt{[DANGER]}
  \hfill \textit{Your papers are in order. The Black Hand does not care. They take you anyway. The order was the trap.}
\item \textbf{Esoiti Audit} --- A magistrate demands to see your records. Your records are lies. \texttt{[LAW]} \texttt{[WEIGHT]}
  \hfill \textit{The magistrate smiles. ``Everyone lies. I just need to know how much.'' His quill is already moving.}
\item \textbf{Spirit-Hound Scent} --- A pack of bound hounds is on your trail. \texttt{[DEATH]} \texttt{[TRACKING]}
  \hfill \textit{The hounds are invisible. You can hear their breathing, close and wet. Their breath smells of brass and old prayers.}
\item \textbf{Slave Raid} --- A Veiled Sister's caravan is taking villagers. You are in the way. \texttt{[COMBAT]} \texttt{[WEIGHT]}
  \hfill \textit{The villagers are armed. They have been expecting you. You are the bait. The Sister is not here. She is watching.}
\item \textbf{Ikasha's Whisper} --- In a quiet moment, you hear a voice: ``Free me.'' \texttt{[OMEN]} \texttt{[PATRON]}
  \hfill \textit{The whisper comes from a crack in the floor. A single green eye looks up. The eye is glass. It blinks.}
\item[J] \textbf{The Ninth Name Spoken} --- Someone in the party speaks a dead name aloud. \texttt{[DEATH]} \texttt{[WEIGHT]}
  \hfill \textit{The air grows cold. A spirit appears. It wants to know why you called. It does not accept silence as an answer.}
\item[Q] \textbf{Sister's Test} --- A Veiled Sister offers a deal: betray a Breath-less contact for safety. \texttt{[SOCIAL]} \texttt{[WEIGHT]}
  \hfill \textit{The offer is genuine. The price is your soul. The Sister watches you decide. She is not in a hurry. She has all of eternity.}
\item[K] \textbf{Mass Haunting} --- The bound spirits rise as one. The Veil is tearing. \texttt{[DEATH]} \texttt{[HORROR]}
  \hfill \textit{The spirits do not attack. They simply ask: ``Why did you let this happen?'' They do not mean the empire. They mean the children.}
\item[A] \textbf{The Demon's Weight Called} --- Varash the Unmaker (or a fragment) demands payment. \texttt{[DEATH]} \texttt{[WEIGHT]}
  \hfill \textit{The amulet Kalrantha glows. A voice whispers: ``Speak your desire.'' The whisper is your own. You do not remember giving it permission.}
\end{enumerate}

\paragraph*{(Name/Lamp/Seal)} A slave's true name (free them or bind them further); spirit-lamp (one question to a bound spirit); brass seal (guest status in a Sister's sanctum).

\section*{\diamondcard{A} Rewards/Leverage}
\label{sec:ashaan-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Slave's True Name} --- Use it to free a bound person — or bind them further. \texttt{[WEIGHT]} \texttt{[POWER]}
  \hfill \textit{The name is written on a scrap of skin. The skin was the slave's own. The letters are still warm.}
\item \textbf{Spirit-Lamp} --- Ask one question of a bound spirit. It cannot lie. \texttt{[MAGIC]} \texttt{[TRUTH]}
  \hfill \textit{The lamp burns green. The spirit's face appears in the flame. It is weeping. The tears evaporate into glass.}
\item \textbf{Brass Seal (Guest Status)} --- Safe passage in a Sister's sanctum for one visit. \texttt{[SOCIAL]} \texttt{[PROTECTION]}
  \hfill \textit{The seal is warm. It glows when a Sister is lying. It also glows when you are lying. The Sister does not ask which.}
\item \textbf{Veil-Clip} --- Speak without being silenced by Maeri magic. \texttt{[MAGIC]} \texttt{[PROTECTION]}
  \hfill \textit{The clip is a small brass spider. It crawls onto your collar and whispers. It whispers the names of those who cannot speak.}
\item \textbf{Erased Name Page} --- A page from the Esoiti record with a name struck through. \texttt{[SECRET]} \texttt{[WEIGHT]}
  \hfill \textit{The ink is still wet. The name was erased ten minutes ago. The page is warm. It has a pulse.}
\item \textbf{Breath-less Cipher} --- A codebook for reading secret messages on water. \texttt{[SECRET]} \texttt{[STEALTH]}
  \hfill \textit{The cipher is written on a leaf. The leaf dissolves when read. The water remembers. The water does not forgive.}
\item \textbf{Spirit-Hound Whistle} --- Call off a pack of bound hounds once. \texttt{[PROTECTION]} \texttt{[DEATH]}
  \hfill \textit{The whistle is made from a child's bone. The child was the hounds' last victim. The hounds remember her scent.}
\item \textbf{Ikasha's Dream} --- One prophetic vision of the Maeri's next move. \texttt{[MAGIC]} \texttt{[KNOWLEDGE]}
  \hfill \textit{The dream arrives as a headache. The vision is clear. The cost is one memory. The memory is always of a door closing.}
\item \textbf{Black Hand Mask} --- Wear it to pass as a Hunter for one scene. \texttt{[STEALTH]} \texttt{[DECEPTION]}
  \hfill \textit{The mask has no eye holes. The mask sees for you. It also remembers. It will report what it saw to the Hand.}
\item[J] \textbf{Cage-Maker's Flaw} --- A key that opens any spirit-cage. One use. \texttt{[MAGIC]} \texttt{[PROTECTION]}
  \hfill \textit{The key is a bent nail. It was the flaw the cage-maker left. The flaw is also a prayer. The cage-maker does not know to whom.}
\item[Q] \textbf{Drowned Sister's Secret} --- The location of a hidden Breath-less archive. \texttt{[SECRET]} \texttt{[KNOWLEDGE]}
  \hfill \textit{The archive is in a well. The well is guarded by a ghost. The ghost's name is Mercy. She does not remember why.}
\item[K] \textbf{Ninth Sister's Veil} --- Wear it to appear as a Veiled Sister for one hour. \texttt{[DECEPTION]} \texttt{[DANGER]}
  \hfill \textit{The veil is also a noose. Do not tie it too tight. The Sister who gave it to you is already dead. She did not untie it.}
\item[A] \textbf{The Unnamed Prince's Map} --- A map to the original Kalrantha temple. \texttt{[LEGEND]} \texttt{[QUEST]}
  \hfill \textit{The map is drawn on human skin. The skin was the prince's own. The map glows under moonlight. It shows the way — and the cost.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:ashaan-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2--5 \(\rightarrow\) 4, 6--10 \(\rightarrow\) 6, J/Q/K \(\rightarrow\) 8, A \(\rightarrow\) 10).
\item Diamonds are codified outcomes (names, lamps, seals) that change position rather than call for a roll.
\item If any A appears, echo \textbf{veil-omens}—brass chains that rattle without wind, scrolls that open to the wrong page, a whisper that sounds like your own voice, green glass growing where no glass should be.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:ashaan-mechanics}

\begin{itemize}
\item \textbf{The Veil:} The Maeri's control is maintained through spirit-binding and terror. The Veil is a magical membrane that separates the living from the dead. When the Veil fails, spirits walk. Mechanically, the GM may spend 3 SB to trigger a \emph{Veil Failure} scene: all spirits become visible, bound spirits may break free, and social rolls against the Maeri's agents gain +1 Position. The Veil fails more often now. No one knows why.

\item \textbf{The Veil's Attention (Divine Threat Clock):} Unlike mundane factions, the Veiled Sisters do not have a favour or debt track. They have a timer: \textbf{Veil's Attention [6]}. It ticks when the party draws the Sisters' gaze — by speaking a forbidden name, by freeing a slave, by resisting a Black Hand sweep, or simply by being in the wrong place at the wrong time. When the timer ticks, the GM may impose an omen: a mirror fogs, a bell rings in still air, a shadow falls where no light is cast. When the timer reaches 6, a Veiled Sister appears. She does not offer a bargain. She pronounces a \emph{judgment}: a name to be erased, a memory to be taken, a life to be claimed. The party cannot fight her. They can only flee, or hide, or offer a substitution — a name they are willing to lose. The Sisters are not enemies. They are a natural disaster with a ledger.

\item \textbf{The Breath-less (Ikasha's Cult as Active Resistance):} She Who Sleeps is not a patron of open rebellion. Her followers meet in root cellars, learn to read records without leaving marks, and whisper secrets under the cover of a cough. A character who whispers Ikasha's name before sleeping gains +1 die to resist fear for the next day, but marks 1 \textbf{Exposure} (the Black Hand may notice). Their symbol is a broken brass chain. Aiding the Breath-less may earn a \textbf{Breath-less Cipher} (see Diamonds) but advances a \textbf{Black Hand Interest} timer.

\item \textbf{The 999 Immortals:} The failed invasion of Oshiira is commemorated in the Temple of the Golden Sepulcher, where the gem-eyed skulls of the 999 are interred. Each New Flood, magistrates process with gold-veiled reliquaries, praying not to repeat their hubris. The guardless helmed blades of the Immortals became trophies; many now hang at the hips of Oshiira's greatest warriors. Some say the skulls whisper at night. They whisper a name that no one will speak.

\item \textbf{Veil Etiquette:} At court, a face fully veiled in coin signals ``I speak as the treasury.'' To address such a speaker is to accept their price. The ninth cup at elite tables is lifted without drinking — a silent remembrance of the 999 and a warning against hubris. Masktime (the hour before sunrise) is when contracts are sealed, lovers dismissed, and bribes counted; unmasked business at that hour is considered offensive. Those who break the ninth cup rule find that their next drink tastes of glass.

\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Unbound Chain} -- Ikasha. \textit{She Who Sleeps; her dream is freedom. Her dream is leaking.}
    \item \textbf{The Gilded Scale} -- Mykkiel. \textit{The Maeri's twisted version of divine law; the zealots of Galanina. Their scales weigh not coin, but names.}
    \item \textbf{The Drowned Prince} -- Khemesh. \textit{Pressure from the deep; the Red Desert was once his seabed. His cult practices child sacrifice in secret. The sacrifices are always the ninth child.}
    \item \textbf{The Silvered Tongue} -- Malachai. \textit{Curses sold as blessings, whispered in the slave markets. The blessing always includes a ninth clause. No one reads it.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Kalrantha Amulet (Dhahara Connection):} The Red Desert was once fertile land, destroyed by an Ashaani prince who stole the Kalrantha amulet. The Maeri want the amulet to bind the demon again — or to use it for themselves. The Breath-less want to destroy it. The prince's true name, long severed by the Pereshi, is said to be written on a scroll sealed beneath Mount Khvarena. The Maeri would pay anything to learn it. The sand remembers his crime. The sand is still hungry.
  
  \item \textbf{The Sidhi of the Brass Coast (Way of Silk):} The Sidhi are coastal cousins of the Ashaani who fled slavery generations ago. They maintain lighthouse courts and freedman's seals. They remember every raid. A Sidhi captain may be a Breath-less contact — and some whisper that the green flame of their lighthouses is a fragment of Ikasha's dream, stolen from the Veiled Sisters. The flame flickers when the Veil tears.
  
  \item \textbf{The Veiled Sisters and the Serpent's Turn (Isoka):} The oldest Veiled Sister legends speak of a time when name-erasure was a sacred judgment, witnessed by the Serpent (Isoka). The severed name was buried in a well of tears, and the Serpent's gaze kept the Hollow at bay. But when the Sisters began erasing names for silver and power, the Serpent turned away. Now the Serpent's gaze falls only on the Breath-less — who whisper that Ikasha's dream is to restore what the Serpent abandoned. The Sisters do not speak of this. The Hollow does not care. The Serpent's shed skin is said to be buried beneath the Brass Gate.
  
  \item \textbf{The Hollow and the Ninth Name (Eastern Realms):} In Ashaan, the Hollow is the space left by an erased name. The Maeri feed the Hollow with forgotten weights. The Breath-less know that to restore a name is to weaken the Veil — and that the Hollow's hunger is the price of the Serpent's silence. A name restored can weaken a Sister's binding for a season. The Hollow is patient. The Hollow is not hungry. The Hollow is \emph{waiting}.
  
  \item \textbf{Oshiira's Cold Peace:} Grain barges from Oshiira pay ruinous tariffs at the Brass Gate. In return, Ashaani spies watch every weigh-house in Mbaro, waiting for the Widow to falter. A party that smuggles grain past the Gate could earn the gratitude of starving villages — and the ire of the Veiled Sisters. The Breath-less sometimes bribe the gatekeepers with stories, not silver; stories are harder to trace. The gatekeepers have a ninth story they never tell.
  
  \item \textbf{Taharka's Unconquered Passes:} Ashaani armies have shattered against the Monsoon Crown's highlands. Now the Maeri send agents disguised as spice merchants to spread poison and dissent. A Breath-less cell in the Silver Pass may hold the only copy of a treaty that could unite Taharka and Ashaan against a common foe — if anyone is brave enough to read it. The treaty is written on a scroll sealed with a knotted cord; the knot is Tulkani work, and it cannot be cut without a story. The ninth page of the treaty is blank. It is reserved for the name of the traitor.
  
  \item \textbf{The Pereshi Highlands and the Forgotten Prince:} The fire temples of Pereshi guard the severed name of the Ashaani prince who tried to conquer them. The Maeri would pay anything to learn that name — for to speak it would be to unmake the Veiled Throne's own claim to authority. The Pereshi will never give it. A party caught between the two might steal the name — or destroy it forever. The Breath-less believe that destroying the name would free the Hollow's attention from Ashaan; the Maeri believe it would let them bind the Hollow itself. The prince's name is carved on a ninth coal that has never been lit. It is warm.
\end{itemize}

\subsection*{Ashaan Superstitions (Burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Veil:} Crumble a crust at a brass cage. Ignore the first \emph{Spirit-Dust Storm} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Veil-Clip:} Tie a single knot in your veil. Cancel \emph{Black Hand Sweep} or take --1 die on your next stealth roll (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Well:} Whisper a dead Breath-less agent's name into a well. Gain +1 Effect when evading the Maeri this scene, but mark \emph{Obligation} to Ikasha. \texttt{[SUPERSTITION]}
  \item \textbf{Glass to the Sand:} Drop a shard of green glass into the Red Desert at noon. The next time the Veil fails, you will see it coming one minute in advance. The glass will hum the warning. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:ashaan-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Black Hand agents or Esoiti clerks}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{spirit-hounds or slave-catchers}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{Veiled Sisters or bound spirits}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or safe passage}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local Black Hand or Esoiti magistrate} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{veil failure or spirit manifestation}.
\item \textbf{Foe Appears:} A \textbf{bound spirit or slave-hunter} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Black Hand, Esoiti guards, or bound spirits} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes — \textbf{Veil fails / spirit breaks free / name is spoken}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{A Maeri (Veiled, Helmed, or Masked), Esoiti High Magistrate, or Ikasha's manifestation} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically — \textbf{Veil tears / sandstorm rises / Brass Gate seals}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice — \textbf{omen appears / blessing lost / curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Ashaan (The Veil):} Brass chains rattle, scrolls open to the wrong page, a whisper sounds like your own voice, a green glass shard grows from a crack in the wall.
\item \textbf{Ashaan (Red Desert):} Glass shards hum, sand shifts to reveal bones, a mirage shows a memory you had forgotten, the sky turns the colour of a healing bruise.
\item \textbf{Ashaan (Breath-less):} A cough from a shadow, a knock in a cellar, a name written on water that does not dissolve, a candle that burns without a wick.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth):} On a hit, shift one band. On a strong hit, also impose \emph{False Trail}.
\item \textbf{Cut the Veil (Presence + Sway):} On a hit, freeze range for one exchange; if a Sister's seal is in your possession, +1 die.
\item \textbf{Weather the Haunting (Resolve + Spirit):} On a hit in \emph{Mass Haunting}, you pick who is targeted; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Spirit \& Veil)}
\label{sec:ashaan-bosses}

\begin{itemize}
\item \textbf{The Record-Eater (Nine Erasures)} — \emph{Phases}: Audit Thresholds \(\rightarrow\) Orchestrated Erasure \(\rightarrow\) Veil Coup. \emph{Cracks}: cut a lineage page; reveal swapped seal; survive a staged haunting. It appears as a scribe whose hands are brass.
\item \textbf{The Unbound Sister (Brine-Veil)} — \emph{Phases}: Drownings Called Mercy \(\rightarrow\) Procession at Brass Gate \(\rightarrow\) Spirit-Baptism of a Magistrate. \emph{Cracks}: relic spirit-lamp, survivor testimony, the Veil turning against them. Her veil is made of chain, not silk. It clinks when she moves.
\item \textbf{The Forgotten Prince's Echo (Pereshi Connection)} — \emph{Phases}: Whisper of the True Name \(\rightarrow\) Fire Temple Cracks \(\rightarrow\) The Prince Walks. \emph{Cracks}: burn a page of the Pereshi epic, speak a different name into the Ninth Coal, or convince the Maeri that the prince was never real. His shadow has no feet. He walks on glass.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Ashaani, power flows through fear, spirit-binding, and the erasure of names. The Maeri hold ultimate authority, but the Esoti control the records, the Black Hand enforces silence, and the Breath-less whisper in cellars. A name can be a weapon or a shield. The Mykkielites of Galanina are an acid slowly eating the old order, and Ikasha stirs in her sleep.

But the true power is the \emph{creep}. No one knows when the Veil will tear next. No one knows which name will be the ninth. No one is safe. Not the Maeri. Not the Breath-less. Not the slaves, not the masters, not the glass that grows in the dark. The punishment is spreading. The sin is forgotten. The sand remembers. The sand is patient.

\textbf{For the GM:}  
Power in Ashaani society is hidden and treacherous. Patronage often takes the form of true names, spirit-lamps, or erased pages — each carrying the weight of damnation. To emphasize this:
\begin{itemize}
\item Tie rewards to secrets (names, whispers, visions) that can be betrayed, stolen, or used against the giver.
\item Let rival factions (Maeri vs. Esoti vs. Breath-less vs. Mykkielites) offer conflicting paths, forcing players to choose whose weight to incur.
\item Use the Brass Gate, the Record Hall, and the root cellars as arenas for social contests, where truth is a weapon and silence is a shield.
\item Track \textbf{Black Hand Interest} as a campaign timer. Each time the party aids the Breath-less or escapes a Sister, tick the timer. When it fills, a Black Hand squad arrives with a warrant for their erasure — but the warrant is signed by a Sister who has been dead for a year.
\item Let the Veil fail unpredictably. The GM may spend SB to trigger a tear at the worst moment. The tear is always in the shape of a ninth door.
\end{itemize}
In Ashaani society, your name is your bond — and your bond can be erased. But the erasure leaves a scar. The scar grows glass.
\end{tcolorbox}

% Index entries
\index{Ashaan|see{Veiled Throne}}
\index{Theme generator}
\index{Veiled Throne generator}
\index{Places!Ashaan}
\index{People!Ashaan}
\index{Complications!Ashaan}
\index{Rewards!Ashaan}
\index{Maeri (Three Sisters)}
\index{Veiled Sister}
\index{Helmed Sister}
\index{Masked Sister}
\index{Esoiti}
\index{Black Hand}
\index{Breath-less}
\index{Ikasha}
\index{Mykkiel}
\index{Red Desert}
\index{Brass Gate}
\index{Galanina}
\index{Khasurah (swallowed city)}
\index{Spirit-binding}
\index{999 Immortals}
\index{Kalrantha amulet}
\index{Green glass}
\index{Local Patrons!Ashaan}