% ============================================================
% Ayokha — The Monsoon Throne
% Revised with Southeast Asian mythology, folklore, and additional generators
% Debt/ledger language reduced to subtext; focus on promise, memory, offering
% ============================================================
\section{Ayokha --- ``The Monsoon Throne''}
\label{chap:ayokha}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The rain writes the law. The river witnesses. The ninth cup is for the drowned.}

\begin{quote}
  \textbf{God-King's Astrologer (Elite):}  
  ``The monsoon does not come because we pray. It comes because the stars align. I read the stars. I align them. Without me, the rains would fail, the rice would burn, and the kingdom would fall. I am not a servant of the God-King. I am his monsoon.''
\end{quote}

\begin{quote}
  \textbf{River Merchant (Commoner):}  
  ``My boat has more secrets than cargo. The lower deck is for silk. The upper deck is for whispers. The tiller is for favors. I have carried nine spies this year. The ninth was the God-King's own daughter.''
\end{quote}

\noindent\textbf{Genre / Mood:} Monsoon noir – waterlogged intrigue, spirit bargains, and the tyranny of the rain.

\noindent\textbf{Starting Location:} A river market on stilts, where boats load silk and spices, and the air smells of brine and betrayal. A Lethai-ar Vigil holds a green lamp: ``The God-King's astrologer has predicted a late monsoon. The merchants are panicking. The rice will fail unless someone makes an offering to the river spirit. You are someone. Go.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Ayokha is the land of the Monsoon Throne, where the God-King rules by the rhythm of the rains. Rivers flood, rice terraces glow green like jade stairways, and the wharves are crowded with ships from Dhahara, Sihai, and distant isles. The Lethai-ar witness every contract; the spirit-mediums speak for the \emph{nats} of the ford – hungry spirits who live in eddies and under bridges. Power here is measured in monsoon schedules, silk bales, and the favor of the God-King's astrologer, a man who has not seen the sun in thirty years. The ninth taboo is absolute: never pour the ninth cup, never count the ninth wave, never speak the ninth name. The Hollow watches from the flooded fields, and the drowned sometimes knock on the hulls of passing boats.

\vspace{2mm}
\textbf{What you notice first:} The smell of wet earth and fermenting rice, thick as a fever dream. The way every bridge has a spirit shrine at its foot, offerings of betel leaf and candle wax. The constant sound of water — dripping, flowing, raining, whispering.

\textbf{The one rule that kills newcomers:} Never refuse a gift of betel leaf. It is an oath of hospitality. Refusing it is an insult that can end a bloodline. And never pour the ninth cup — the ninth is for the river, and the river is always thirsty.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Monsoon}

Power here is measured in monsoon predictions, silk bales, and the number of spirits placated. The God-King rules from a floating palace, but the astrologer sets the calendar, the Lethai-ar witness the contracts, and the spirit-mediums negotiate with the nats. A merchant who ignores a nat's demand will find his boat sunk and his name struck from the wharf records. A farmer who disrespects the monsoon will find his fields barren and his children dreaming of drowning. In Ayokha, the rain is not weather – it is a \emph{promise}, and every promise has a season.

\subsection*{Prominent NPCs of Ayokha}
\label{sec:ayokha-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
God-King's Astrologer & Royal astronomer & Predict the monsoon before the people starve & He has not seen the sun in thirty years. He works by candlelight. \\
\hline
Vigil Lady Suyin & Lethai-ar witness-master & Keep the river's records free of forgeries & She has cut nine witness cords. The ninth was her own. \\
\hline
River Merchant & Silk trader & Smuggle contraband under the God-King's tariffs & She has a hidden deck that only appears at low tide. \\
\hline
Spirit-Medium & Nat negotiator & Speak for the spirit of the ford & She carries a bowl of water; the spirit speaks through ripples. \\
\hline
The Drowned Nat (Ghost) & A spirit of a broken oath & Demand payment before the monsoon fails & It appears as a wet figure in torn robes. It never blinks. \\
\hline
\end{longtable}

\subsection*{Names of Ayokha}
\label{sec:ayokha-names}

Inspired by the monsoon kingdoms: soft, flowing consonants, honorifics (Lady, Master). Surnames denote a river, a wharf, or a spirit.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/Epithet} & \textbf{Monsoon-Name} \\
\hline
Hiro & Suyin & River-watch & \textit{the Uncut} \\
\hline
Tama & Mira & Silk-hand & \textit{Ninth Knot} \\
\hline
Kaito & Yuki & Lamp-bearer & \textit{Green Flame} \\
\hline
Riku & Hana & Oath-strand & \textit{the Witness} \\
\hline
Sora & Nami & Mud-dyer & \textit{the Unforgotten} \\
\hline
\end{longtable}

\paragraph*{(Market/Temple/Bridge)} A river market on stilts; a step-pyramid temple with monsoon carvings; a bridge that appears only once a century.

\section*{Spades — Places}
\label{sec:ayokha-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{River Market (Stilts)} --- A floating market where goods are passed from boat to boat. `[COMMERCE]` `[SOCIAL]`
  \hfill \textit{The prices are written in water. They change with the tide.}
\item \textbf{Step-Pyramid Temple} --- A stone pyramid rising from the riverbank. Carvings of the first monsoon. `[RELIGION]` `[WEATHER]`
  \hfill \textit{The carvings weep water during the rainy season. The water is holy.}
\item \textbf{Bridge of Three Echoes} --- A wooden bridge that appears only once a century. `[MYSTERY]` `[THRESHOLD]`
  \hfill \textit{Those who cross it see three versions of themselves: past, present, and future.}
\item \textbf{Silk Village (Mulberry Groves)} --- A village where silk is spun and cords are woven. `[CRAFT]` `[OATH]`
  \hfill \textit{The mulberry trees were planted by the first God-King. Their leaves are sacred.}
\item \textbf{Astrologer's Observatory} --- A roofless tower with a bronze armillary sphere. `[MAGIC]` `[SECRET]`
  \hfill \textit{The sphere has nine rings. The ninth ring is broken. The astrologer refuses to fix it.}
\item \textbf{River Court (Floating)} --- A barge that serves as a mobile courthouse. `[LAW]` `[SOCIAL]`
  \hfill \textit{The barge is anchored at a different spot each day. The river is the judge.}
\item \textbf{Spirit-Medium's Hut} --- A small shelter on the riverbank, surrounded by offering bowls. `[SPIRIT]` `[MYSTIC]`
  \hfill \textit{The bowls are filled with rice, flowers, and small coins. The spirits take the offerings at night.}
\item \textbf{Well of the Drowned Nat} --- A stone well where a spirit was trapped. `[DEATH]` `[TOLL]`
  \hfill \textit{The well is covered by a stone slab. The slab has nine holes. The spirit breathes through them.}
\item \textbf{God-King's Palace (Floating)} --- A wooden palace built on pontoons, anchored in the river. `[AUTHORITY]` `[WEALTH]`
  \hfill \textit{The palace rises and falls with the flood. The God-King never touches dry land.}
\item[J] \textbf{The Ninth Terrace} --- A rice terrace that is flooded only once every nine years. `[FORBIDDEN]` `[LEGEND]`
  \hfill \textit{The rice grown there is for the God-King alone. Eating it is treason.}
\item[Q] \textbf{Monsoon Bell Tower} --- A stone tower with a single brass bell. The bell is rung at the first rain. `[WEATHER]` `[RITUAL]`
  \hfill \textit{The bell's tone determines the harvest. A high tone means abundance. A low tone means famine.}
\item[K] \textbf{The First Bridge (Myth)} --- Where the first God-King crossed the river and claimed the throne. `[LEGEND]` `[THRESHOLD]`
  \hfill \textit{The bridge is invisible. Only those who have witnessed nine true oaths can see it.}
\item[A] \textbf{The Monsoon's Eye} --- A still pool in the astrologer's observatory. It reflects the sky. `[OMEN]` `[MAGIC]`
  \hfill \textit{The pool shows the future of the monsoon. The astrologer is the only one who can read it.}
\end{enumerate}

\paragraph*{(Merchant/Medium/Vigil)} River merchant with a boat of secrets; spirit-medium with a bowl of water; Lethai-ar Vigil with a green lamp.

\section*{Hearts — People \& Factions}
\label{sec:ayokha-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{River Merchant} --- A woman with a brass earring and a hidden deck. She never sails without a bribe. `[COMMERCE]` `[DECEPTION]`
  \hfill \textit{``The tariff is negotiable,'' she says. ``My silence is not.''}
\item \textbf{Spirit-Medium} --- A woman with a bowl of water. She watches the ripples. `[SPIRIT]` `[SOCIAL]`
  \hfill \textit{``The nat says the river is angry,'' she says. ``Someone broke a promise.''}
\item \textbf{Vigil Lady Suyin} --- Gray-robed, a green lamp at her hip. She never blinks when witnessing an oath. `[LAW]` `[OATH]`
  \hfill \textit{``The cord will tighten,'' she says. ``I do not need to watch. The river watches for me.''}
\item \textbf{God-King's Astrologer} --- A pale man in silk robes, surrounded by candles. `[MAGIC]` `[SECRET]`
  \hfill \textit{``The monsoon will be late,'' he says. ``The rice will fail unless...'' He trails off.}
\item \textbf{Rice Farmer} --- A woman with mud-stained feet and a straw hat. She leaves offerings at the terrace edge. `[SURVIVAL]` `[SPIRIT]`
  \hfill \textit{``The spirits are hungry,'' she says. ``Feed them, or they feed on you.''}
\item \textbf{Silk Weaver} --- A man with thread between his fingers. He weaves witness cords for the Lethai-ar. `[CRAFT]` `[OATH]`
  \hfill \textit{``The cord must be perfect,'' he says. ``A single flaw can break an oath.''}
\item \textbf{Monsoon Bell Ringer} --- A young woman who rings the bell at the first rain. She has never been late. `[RITUAL]` `[WEATHER]`
  \hfill \textit{``The rain waits for me,'' she says. ``I do not wait for the rain.''}
\item \textbf{Drowned Nat's Ghost} --- A wet figure in torn robes, standing by the well. `[DEATH]` `[TOLL]`
  \hfill \textit{It asks: ``Who sealed me here?'' The answer is a name. The name is the God-King's.}
\item \textbf{God-King (Offstage)} --- Never seen, rarely heard. His decrees are delivered by astrologer. `[AUTHORITY]` `[MYSTERY]`
  \hfill \textit{Rumor says he is a spirit. Rumor says he is a child. Rumor says he is the river itself.}
\item[J] \textbf{The First God-King's Ghost} --- A crowned skeleton in the palace's lower deck. `[DEATH]` `[LEGEND]`
  \hfill \textit{He asks: ``Is the monsoon still on time?'' If you say no, he weeps.}
\item[Q] \textbf{Monsoon Spirit (Kami)} --- A whirlwind with eyes, seen over the river during storms. `[SPIRIT]` `[WEATHER]`
  \hfill \textit{It asks: ``Who rang the bell?'' The answer must be the ringer's name.}
\item[K] \textbf{The Ninth Terrace's Keeper} --- An old woman who tends the forbidden rice. `[FORBIDDEN]` `[WISDOM]`
  \hfill \textit{She has never eaten the rice. She has never been tempted.}
\item[A] \textbf{Kuva's Shadow (Hollow)} --- A vast darkness that covers the river at midnight. `[DEATH]` `[HORROR]`
  \hfill \textit{It asks: ``Who poured the ninth cup?'' The answer is a name. The name is the speaker's.}
\end{enumerate}

\paragraph*{(Flood/Rumor/Offense)} The river floods — a bridge is swept away; a rival merchant spreads rumors of cursed silk; the nat of the well is offended — no one may draw water.

\section*{Clubs — Complications/Threats}
\label{sec:ayokha-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{River Floods} --- The monsoon rises. The bridge is gone. The wharf is isolated. `[WEATHER]` `[DISASTER]`
  \hfill \textit{The flood is not natural. A river-spirit is angry. An oath was broken upstream.}
\item \textbf{Rumor of Cursed Silk} --- A merchant claims that a bale of silk is cursed. Trade stops. `[SOCIAL]` `[DECEPTION]`
  \hfill \textit{The curse is a lie. The merchant wants to drive down prices.}
\item \textbf{Well-Spirit Offended} --- A traveler threw trash into the sacred well. The spirit refuses water. `[SPIRIT]` `[TOLL]`
  \hfill \textit{The spirit demands an apology. The apology must be an oath. The oath must be witnessed.}
\item \textbf{Monsoon Late} --- The astrologer's prediction fails. The rains do not come. `[WEATHER]` `[CRISIS]`
  \hfill \textit{The astrologer is a fraud. Or the spirits are angry. Either way, the rice is dying.}
\item \textbf{Witness Cord Cut (False)} --- Someone cuts a witness cord that was tied in good faith. `[OATH]` `[BETRAYAL]`
  \hfill \textit{The cutter is the Vigil's own apprentice. She thought the oath was false. It was true.}
\item \textbf{Silk Village Poisoned} --- Someone poisons the mulberry trees. The silkworms die. `[CRIME]` `[ECONOMY]`
  \hfill \textit{The poisoner is a rival merchant. He wants to monopolize the silk trade.}
\item \textbf{River-Spirit Demands a Story} --- The nat appears at the ford. It asks for a broken oath. `[SPIRIT]` `[TOLL]`
  \hfill \textit{It will not leave until someone confesses. The confession must be true.}
\item \textbf{God-King's False Oath} --- The king swears a vow he does not intend to keep. The Vigil knows. `[LAW]` `[POLITICS]`
  \hfill \textit{If she exposes him, the king will exile the Lethai-ar. If she stays silent, the cord tightens.}
\item \textbf{Astrologer's Betrayal} --- The astrologer sells the monsoon schedule to a foreign power. `[TREASON]` `[WAR]`
  \hfill \textit{The foreign power is Ashaan. They want to time an invasion with the floods.}
\item[J] \textbf{The Ninth Cup Poured (By Mistake)} --- A stranger pours the ninth cup at a river market. The Hollow comes. `[DEATH]` `[TABOO]`
  \hfill \textit{The Hollow asks: ``Who poured this cup?'' The stranger points at the party.}
\item[Q] \textbf{First Bridge Appears} --- The mythic bridge manifests. Crossing it grants a new name. `[THRESHOLD]` `[TOLL]`
  \hfill \textit{The new name is a promise. The promise is to the first God-King. He will collect.}
\item[K] \textbf{Monsoon Bell Cracked} --- The bell is struck too hard. It cracks. The tone is wrong. `[SABOTAGE]` `[OMEN]`
  \hfill \textit{The bell-ringer is a spy. She cracked it on purpose. The monsoon will fail.}
\item[A] \textbf{The Monsoon Fails} --- The rains do not come for nine weeks. The river dries. The Hollow walks. `[APOCALYPSE]` `[DEATH]`
  \hfill \textit{The Hollow asks: ``Who failed to ring the bell?'' The answer is a name.}
\end{enumerate}

\paragraph*{(Schedule/Voucher/Cord)} Monsoon schedule (predicts safe sailing days); silk voucher (trade goods at 10\% profit); Lethai-ar witness cord (enforce any oath).

\section*{Diamonds — Treasures \& Discoveries}
\label{sec:ayokha-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Monsoon Schedule (Astrologer's Copy)} --- A scroll of safe sailing days for the next season. `[TRAVEL]` `[WEATHER]`
  \hfill \textit{The schedule is written on dried palm leaf. The leaf will crumble after use.}
\item \textbf{Silk Voucher (Merchant's Seal)} --- A stamped document. Trade goods at 10\% profit in any Ayokha port. `[COMMERCE]` `[WEALTH]`
  \hfill \textit{The seal is a fish. The fish glows when the voucher is genuine.}
\item \textbf{Lethai-ar Witness Cord} --- A silk cord dyed with river mud. Enforce any oath. `[OATH]` `[MAGIC]`
  \hfill \textit{The cord tightens when the oath is broken. It will not stop until blood is drawn.}
\item \textbf{God-King's Blessing (Scroll)} --- A royal decree. One national favor in Ayokha. `[AUTHORITY]` `[SOCIAL]`
  \hfill \textit{The scroll is sealed with river clay. The clay is from the palace's foundation.}
\item \textbf{River Merchant's Hidden Deck Key} --- A brass key. Access any smuggler's hold on the river. `[STEALTH]` `[CRIME]`
  \hfill \textit{The key is warm. It glows when a customs boat is near.}
\item \textbf{Spirit-Medium's Bowl} --- A clay bowl. Fill it with water to speak with any river-spirit once. `[SPIRIT]` `[KNOWLEDGE]`
  \hfill \textit{The spirit will answer one question. The answer will be a ripple.}
\item \textbf{Vigil's Lamp (Green Flame)} --- A small oil lamp. Gain +1 die to social rolls in any Lethai-ar court. `[SOCIAL]` `[TRUTH]`
  \hfill \textit{The flame flickers when someone lies. It also flickers when you lie.}
\item \textbf{Silk Weaver's Needle} --- A needle carved from a mulberry branch. Tie a witness cord without cutting the silk. `[CRAFT]` `[OATH]`
  \hfill \textit{The needle can also cut a cord without making a sound. Use once.}
\item \textbf{Monsoon Bell Tone (Recording)} --- A clay cylinder that captured the bell's ring. Play it to summon a monsoon spirit. `[MAGIC]` `[WEATHER]`
  \hfill \textit{The spirit will come. It will demand an offering. The offering must be a promise.}
\item[J] \textbf{First Bridge Plank} --- A piece of the mythic bridge. Cross any river as if on solid ground once. `[TRAVEL]` `[LEGEND]`
  \hfill \textit{The plank crumbles after use. The first God-King will remember.}
\item[Q] \textbf{Ninth Terrace Rice (Grain)} --- A single grain from the forbidden terrace. Eat it to gain a vision of the future. `[MAGIC]` `[OMEN]`
  \hfill \textit{The vision is always a flood. The flood is always coming.}
\item[K] \textbf{Astrologer's Sphere Ring (Broken)} --- The ninth ring from the armillary sphere. Touch it to predict the monsoon's arrival. `[WEATHER]` `[MAGIC]`
  \hfill \textit{The ring is cold. It warms when the rains are nine days away.}
\item[A] \textbf{The Monsoon's Eye (Water)} --- A vial of water from the astrologer's pool. Drink it to see the future of the harvest. `[VISION]` `[TOLL]`
  \hfill \textit{The water tastes of salt. The vision is always a famine. The famine is not inevitable.}
\end{enumerate}

% ============================================================
% ADDITIONAL GENERATORS (Southeast Asian Mythology & Folklore)
% ============================================================

\subsection*{Monsoon-Spirit Negotiation (d6 Outcomes)}
\label{sec:ayokha-monsoon-spirit}

When the party attempts to bargain with a monsoon spirit (at a shrine, during a drought, or at the Monsoon Bell Tower), roll on this table to determine the outcome. The GM sets the initial DV based on the offering (a silk voucher, a story, or a promise witnessed by a Lethai-ar).

\begin{longtable}{|p{2cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{Outcome} & \textbf{Result} \\
\hline
\textbf{Critical Success} & The spirit grants a \textit{Monsoon Schedule} (see Diamonds) or advances the Monsoon Clock by –2 (rains arrive early). The spirit may also whisper the location of a hidden spring or a sunken relic. \\
\hline
\textbf{Success} & The spirit delays the drought. The Monsoon Clock does not advance for 1d4 days. Gain +1 \textit{River Favor}. \\
\hline
\textbf{Partial} & The spirit accepts the offering but demands an additional gesture: a broken oath must be confessed, or a ninth cup poured into the river. Gain an \textit{Obligation Clock} [4]. \\
\hline
\textbf{Failure} & The spirit is offended. The Monsoon Clock advances by +2. All river travel is Desperate for the next leg. The spirit may send a \textit{nat} to harass the party (see Nat Encounters table). \\
\hline
\textbf{Critical Failure} & The spirit calls the ninth wave. A flash flood sweeps the area. Test Body + Athletics (DV 5) or be swept away. Lose 1 Supply and gain the \textit{Marked by the Monsoon} condition (all weather hazards are one step worse for the next journey). \\
\hline
\end{longtable}

\subsection*{Nat Encounters (d8)}
\label{sec:ayokha-nat-encounters}

When the party travels along rivers, crosses bridges, or disturbs spirit shrines, roll on this table for a nat (river-spirit) encounter. Nats are not always hostile — they are keepers of what is owed.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d8} & \textbf{Encounter} \\
\hline
1 & \textbf{The Toll Nat} — Appears at a ford, demanding a betel leaf or a coin. Refusal capsizes the nearest boat. Payment grants safe passage for one day. \\
\hline
2 & \textbf{The Drowned Child} — A child's ghost sits on a rock, weeping. It has lost its name. If the party returns it (by speaking the child's true name from a village record), the nat grants a \textit{River Charm} (ignore one flood hazard). \\
\hline
3 & \textbf{The Serpent Nat (Naga)} — A serpent with golden scales rises from the water. It demands a story about a broken promise. If the story pleases it, the nat blesses the party's boat (+1 Speed for one leg). If the story offends it, the nat capsizes the boat. \\
\hline
4 & \textbf{The Bridge Nat} — A spirit lives under every bridge. It asks: ``Why do you cross?'' Answer truthfully (e.g., ``to trade,'' ``to flee,'' ``to honor a promise'') and the bridge holds. Lie, and the planks splinter. \\
\hline
5 & \textbf{The Ferryman's Nat} — A hooded figure rows a spectral boat. It offers passage to the other side — but the price is a memory of a living person. Accept and lose a random memory (GM chooses). Refuse and the nat follows the party for 1d4 days, appearing in reflections. \\
\hline
6 & \textbf{The Well Nat (Drowned)} — A hand reaches out of a well. It holds a wet scroll. The scroll contains a forgotten water right. Taking it angers the nat (must pay a story or a cup of blood). Leaving it earns the nat's gratitude (+1 \textit{River Favor}). \\
\hline
7 & \textbf{The Laughing Nat} — A trickster spirit that swaps cargo between boats. Roll Wits + Navigation (DV 4) to notice. On failure, the party loses 1 Supply (stolen) and gains 1 random item (a silk scarf, a jar of pickled fish, a forged customs seal). \\
\hline
8 & \textbf{The Ninth Nat (The Hollow's Breath)} — A figure made of mist and rain. It asks: ``Who poured the ninth cup?'' The party must answer with a name or a story. If they cannot, the Hollow gains 1 tick on its Attention clock. If they answer with a true regret, the nat vanishes and the party gains +1 die on their next river roll. \\
\hline
\end{longtable}

\subsection*{Rice Terrace Hazards (d6)}
\label{sec:ayokha-rice-hazards}

When the party travels through the flooded rice terraces (especially during the monsoon), roll on this table for environmental hazards.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d6} & \textbf{Hazard} \\
\hline
1 & \textbf{Levee Breach} — Water pours through a crack. Test Body + Athletics (DV 4) to avoid being swept down the terrace. Failure costs 1 Fatigue and 1 Supply (rice washed away). \\
\hline
2 & \textbf{Hungry Ghosts (Farmers)} — The spirits of farmers who starved during a drought appear, demanding a share of the harvest. Offer 1 Supply or face a ghost attack (Cap 2, Harm 1). \\
\hline
3 & \textbf{Mudslide} — A section of the terrace collapses. Navigation becomes Desperate. Lose 1 segment of progress. \\
\hline
4 & \textbf{Crab Swarm} — Thousands of small crabs climb the levees, snapping at ankles. Test Resolve (DV 3) or mark 1 Fatigue. The crabs also eat exposed food — lose 1 Supply if the party fails a Wits + Survival test. \\
\hline
5 & \textbf{Sunken Shrine} — A submerged spirit shrine emerges at low water. Investigating reveals a \textit{Ninth Terrace Rice} grain (see Diamonds) but angers the local nat. The nat will demand a story or a promise. \\
\hline
6 & \textbf{The Ninth Terrace Opens} — The forbidden terrace floods unexpectedly. The party can harvest a single grain of the sacred rice (requiring three actions, each triggering a Hungry Ghost attack). If they succeed, they gain a \textit{Ninth Terrace Rice} grain — but the God-King's astrologer will sense its theft. \\
\hline
\end{longtable}

\subsection*{Monsoon Omens (d8)}
\label{sec:ayokha-monsoon-omens}

When the party consults the astrologer, reads the sky, or experiences a shift in weather, roll on this table for a prophetic omen that affects upcoming travel or social encounters.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d8} & \textbf{Omen} \\
\hline
1 & The river runs red at dawn — a blood promise is overdue. The party's next negotiation with a merchant suffers –1 Position. \\
\hline
2 & The Monsoon Bell rings twice — the astrologer has lied about the schedule. The next weather prediction is automatically wrong. \\
\hline
3 & The Bridge of Three Echoes appears a day early — the Hollow is restless. The first person to cross must test Resolve (DV 4) or gain the \textit{Haunted} condition (nightmares). \\
\hline
4 & A nat asks for a story at the ford. Refusal means the river rises. Accept and tell a true story; the GM gains 1 SB but the party gains +1 \textit{River Favor}. \\
\hline
5 & The Vigil's green lamp burns blue — a witness cord has been cut. The party will encounter a Lethai-ar seeking justice in the next 24 hours. \\
\hline
6 & Rice terraces flood without rain — the God-King has forgotten an offering. The party must perform a \textit{Rain Offering} (costs 1 Supply) or the Monsoon Clock advances by +1. \\
\hline
7 & The Ninth Terrace blooms in winter — the forbidden rice is ready. The keeper is dead. The party can attempt to harvest (see Rice Terrace Hazards) but the astrologer will send guards. \\
\hline
8 & The Monsoon's Eye shows a face — the face is yours, but older. The astrologer will not explain. Gain a \textit{Prophecy Token}: once, reroll a failed roll related to weather or rivers, but the GM may introduce a complication. \\
\hline
\end{longtable}

\section*{Quick use notes}
\label{sec:ayokha-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = treasure. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Treasures (Diamonds) are codified outcomes (schedules, vouchers, cords) that change position rather than call for a roll.
\item If any A appears, echo \textbf{monsoon-omens} — water that flows upstream, a lamp that burns green without oil, a cord that ties itself.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:ayokha-features}

\begin{itemize}
\item \textbf{Rain's Due (Shared Clock, size 8):} Every broken promise to a river-spirit, every unpaid offering, every late monsoon gesture ticks the clock. The party inherits the God-King's obligations at the start of any adventure set in Ayokha – they begin with 2 ticks already accrued (old burdens). When the clock reaches 4: the astrologer's predictions become unreliable (roll twice for weather, take worse). When it reaches 8: a \textbf{Oath-Keeper} rises from the river and demands a story, a memory, or a season of service. Resetting the clock requires a \textbf{Monsoon Rite} (downtime, costs 1 XP or a story worth a season's wages) or a sacrifice to the Drowned Nat (a memory or a magical item).

\item \textbf{Monsoon Clock [6]:} Tracks the approach of the rainy season. When it fills, the monsoon arrives — all river travel becomes Desperate for 2 legs. The clock ticks each day during the dry season. The astrologer can predict the exact day; a successful Wits + Lore test (DV 4) reveals the remaining ticks. The party can slow the clock by performing \textbf{Rain Offerings} (betel leaves, silk scraps, or stories) – each offering ticks the clock back by 1, but adds 1 tick to \textbf{Rain's Due}.

\item \textbf{River Court Justice:} Disputes on the river are settled by a floating court. The judge is a Lethai-ar Vigil. The verdict is written on a clay tablet and thrown into the water. If the tablet floats, the verdict is true. If it sinks, the verdict is false — and the judge is drowned. A character may attempt to influence the verdict by offering a bribe (a Treasure or a story) before the tablet is thrown. If the bribe is a story, the GM gains 1 SB; if a Treasure, the judge's verdict tilts in your favor automatically (no test).

\item \textbf{Betel Leaf Hospitality:} Offering a betel leaf is an oath of hospitality. Refusing is a grave insult. Accepting binds both parties to non-violence for the duration of the visit. Breaking this oath marks 1 Corruption and begins a \textbf{River's Grudge [4]} clock. When the clock fills, a nat appears and demands an equal price (1 Harm, a memory, or a public apology).

\item \textbf{Astrologer's Scroll (Player Resource):} A character who earns the astrologer's trust (or steals his notes) may carry a copy of his calendar. The scroll has 3 pages. Each page can be used to:
  \begin{itemize}
    \item \textbf{Predict the Monsoon:} Once per journey, ask the GM if the weather will turn against you in the next leg. The GM answers truthfully (yes/no).
    \item \textbf{Forge a Schedule:} Once per session, treat a failed navigation roll as a success, but add 1 tick to \textbf{Rain's Due} (the astrologer will notice the forgery eventually).
    \item \textbf{Blackmail an Official:} Gain +1 Position on social rolls against any Ayokha official who relies on the monsoon schedule (harbor masters, customs officers, grain factors). Using this burns the page.
  \end{itemize}
  Refilling the scroll requires a downtime action and a bribe to the astrologer (1 Boon or a story worth a season's gossip).

\item \textbf{The Ninth Terrace (Forbidden Zone, Mechanic):} The terrace is flooded only once every nine years. The rice grown there is for the God-King alone – eating it is treason. But the rice is also the \emph{only} thing that can cure the monsoon blight (a crop disease that ticks the Monsoon Clock by 2 each day). Entering the Ninth Terrace requires a \textbf{Water-Key} (a coral token, acquired from the astrologer or a nat). The terrace is guarded by \textbf{Hungry Ghosts} – the spirits of farmers who starved when the monsoon failed. They attack anyone whose \textbf{Rain's Due} clock is 4 or higher. Harvesting the rice takes 3 actions. Each action triggers a ghost attack (Resolve test DV 4 or take 1 Fatigue). Success yields a \textbf{Ninth Terrace Rice} grain (see Diamonds).

\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Rain That Waits} – Ikasha. \textit{She Who Sleeps beneath the flood; her dream is the monsoon's hesitation.}
    \item \textbf{The Unbroken Cord} – Inaea. \textit{The river's memory woven into silk; every oath is a knot.}
    \item \textbf{The Drowned King} – Khemesh. \textit{Pressure from the deep; the ninth wave is his breath.}
    \item \textbf{The Green Lamp} – The Witness. \textit{The Vigil's flame; it flickers when a lie enters the record.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Kalrantha Fragment (Malachai, the Chained Angel):} A silver coin that never spends was found floating in the Monsoon's Eye. The astrologer keeps it in a jar of brine. It whispers to him at night — promises of perfect predictions, of rains that never fail. The coin's warmth spreads through the palace like a fever.
  \item \textbf{The Drowned Courier (The Grey Wanderer's Grimoire):} A Breath-less agent drowned in the river last monsoon, but her ghost still carries messages. She appears at midnight on the Bridge of Three Echoes, offering to deliver secrets to the dead. The price is a memory of living sunlight.
  \item \textbf{The Root That Drinks (The Bitter Root):} A Daughter of the Cord has taken refuge in the well of the Drowned Nat. She claims the spirit is not a nat but a Hollowed — one of the High Mother's own, lost when a witness cord was cut. She asks the party to help her re-tie the cord before the monsoon fails.
  \item \textbf{The Ninth Terrace's Promise (Ukwe – Sekogo):} The forbidden rice of the Ninth Terrace is said to be grown from seeds stolen from the Ukwe's roots. The Green Council demands its return, but the God-King refuses. A Sekogo druid offers the party a way to steal the rice without angering the spirits — if they also bring back a vial of water from the Monsoon's Eye.
\end{itemize}

\subsection*{Ayokha Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the River:} Crumble a crust into the current. Ignore the first \emph{River Floods} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Cord:} Tie a single knot in a witness cord before using it. Cancel \emph{Witness Cord Cut} or take –1 die on your next oath. `[SUPERSTITION]`
  \item \textbf{Name to the Well:} Whisper a dead Vigil's name into the well. Gain +1 Effect when witnessing an oath, but mark \emph{Obligation} to the Monsoon Spirit. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:ayokha-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{river merchants or spirit-mediums}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{bell-ringer or customs boats}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{river-spirits or the Vigil}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local Vigil or spirit-medium} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{flood or fog}.
\item \textbf{Foe Appears:} A \textbf{rival merchant's boat or a river-spirit} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{spirit-mediums, customs boats, or river-spirits} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes — \textbf{monsoon fails / well runs dry / cord is cut}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{God-King's Astrologer, Vigil Lady Suyin, or the Drowned Nat} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically — \textbf{flood sweeps the market / bell tower cracks / ninth terrace floods}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice — \textbf{omen appears / blessing lost / curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Ayokha (Monsoon Law):} Rain falls upward, the river changes course, a cord ties itself.
\item \textbf{Ayokha (Spirit Law):} A nat laughs from the well, an offering steams, the Vigil's lamp burns blue.
\item \textbf{Ayokha (Rice \& Silk):} A terrace floods for no reason, a silk bale weeps water, the bell rings twice.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Navigation):} On a hit, shift one band. On a strong hit, also impose \emph{Monsoon Swell}.
\item \textbf{Cut the Cord (Presence + Lore):} On a hit, freeze range for one exchange; if a witness cord is in your possession, +1 die.
\item \textbf{Weather the Flood (Resolve + Survival):} On a hit in \emph{River Floods}, you pick who keeps their boat; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Monsoon \& Promise)}
\begin{itemize}
\item \textbf{The Rain-Eater (Nine Dry Fields)} — \emph{Phases}: Wells Sinking $\rightarrow$ Orchestrated Drought $\rightarrow$ Terrace Coup. \emph{Cracks}: find a hidden spring, expose a false monsoon schedule, survive a night on the Ninth Terrace.
\item \textbf{The Drowned Vigil (Cord of the River)} — \emph{Phases}: Broken Oaths $\rightarrow$ Procession of Empty Barges $\rightarrow$ Flood-Baptism of a Merchant. \emph{Cracks}: relic witness cord, survivor testimony, the bell ringer refusing to strike.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In Ayokha, power flows through the monsoon, the river, and the weight of a witnessed oath. The God-King rules from his floating palace, but the astrologer sets the calendar, the Lethai-ar hold the records, and the spirit-mediums speak for the nats. No one rules alone — and everyone owes a tithe to the rain.

\textbf{For the GM:}  
Use the \textbf{Monsoon Clock} as a visible campaign timer — it ticks down to famine or flood. Let \textbf{River Court Justice} be a high-stakes social encounter where a single lie can drown a judge. Enforce \textbf{Betel Leaf Hospitality} as a binding ritual — breaking it invites the Hollow and advances the \textbf{River's Grudge} clock. Offer Treasures as monsoon schedules, silk vouchers, and witness cords — each a tool to survive the next storm or the next audit. Track \textbf{Rain's Due} as the region's central pressure; the party's actions can reduce it (by performing rites or paying sacrifices) or increase it (by breaking oaths or ignoring nats). The God-King's astrologer is a potential ally, but his calendar is for sale to the highest bidder.

The new generator tables — \textbf{Monsoon-Spirit Negotiation}, \textbf{Nat Encounters}, \textbf{Rice Terrace Hazards}, and \textbf{Monsoon Omens} — add depth to supernatural encounters and environmental storytelling, drawing inspiration from Southeast Asian mythologies of river spirits, naga serpents, and the sacred balance of rain and rice.
\end{tcolorbox}

% Index entries
\index{Ayokha|see{Monsoon Throne}}
\index{Theme generator}
\index{Monsoon Throne generator}
\index{Places!Ayokha}
\index{People!Ayokha}
\index{Complications!Ayokha}
\index{Rewards!Ayokha}
\index{Monsoon Clock}
\index{River Court}
\index{Betel Leaf}
\index{Vigil Lady Suyin}
\index{Drowned Nat}
\index{Ninth Terrace}
\index{Monsoon Bell}
\index{First Bridge}
\index{Astrologer}
\index{Rain's Due (clock)}
\index{Astrologer's Scroll}
\index{Local Patrons!Ayokha}
