\section{Black Banners --- Condotta \& Crowns}
\label{chap:black-banners}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{A contract is only as good as the next settlement --- but loyalty is earned in blood, not silver.}

\begin{quote}
  \textbf{Banner-Captain (Elite):}  
  ``A contract is only as good as the next payday, and a grudge is only as sharp as the sword backing it. In this game, we sell steel by the season and vendettas by the year. But a captain who forgets the men in the mud won't last the winter.''
\end{quote}

\begin{quote}
  \textbf{Veteran Sergeant (Commoner):}  
  ``I've seen three captains rise and fall in the time it takes to grow a beard. The mud stays the same. The pay comes when it pleases. But the men? They remember who shared their fire and who ate alone. Keep your blade sharp and your word sharper. The camp sings? You live. They stop singing? Even the best charter bleeds.''
\end{quote}

\noindent\textbf{Genre / Mood:} Mercenary realism, winter warfare, logistics-driven violence, loyalty in the mud.

\noindent\textbf{System Type:} Portable overlay --- apply to any region (Acasia, Violet Steppe, Ecktorian borderlands). Not a fixed geography.

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Condotta \& Crowns}]
Ecktorian treasuries bankroll a perpetual border-war they claim to be \emph{containing}. In truth, the gold keeps the fire burning. The Black Banners are mercenary leagues --- ex-Utaran legionaries, second-sons, deserters, and exiles --- who sell steel to the highest bidder, today's ally becoming tomorrow's quarry. Between them move the Ykrul steppe-clans (swift, pragmatic, oath-canny) and the Vilikari hill tribes (crafty, patient, and proud). Ecktorian ministers call it ``pressure management.'' The companies call it a season --- and a season is measured in the rhythm of pay, but also in snowdrifts, broken wheels, and the weight of a dead friend's shield.

\vspace{2mm}
\textbf{What you notice first:} The smell of wet wool and old blood. The way every tent has a roster pinned to its flap --- and beside it, a list of the fallen. The silence when the chest is opened --- and the murmur when it's light. The way veterans watch the horizon not for enemies, but for the supply train.

\textbf{The one rule that kills newcomers:} Never break a Bell Dawn. Three bells at sunrise means bury the dead, trade prisoners, fix the wells, no ambushes. Violating it is how legends end badly --- and how captains end hanging. The dead don't forget.
\end{tcolorbox}

\begin{tcolorbox}[colback=black!3,colframe=red!60!black,title={The Ninth Taboo Manifestation}]
In any mercenary camp, the ninth tent in a row collapses before dawn. The ground beneath is always damp, as if something rose from below and left. Those who sleep in that tent wake with no memory of the night --- but a new name carved into their belt buckle. The name is never theirs.
\end{tcolorbox}

\subsection*{The Theater of Banners}

Here, war is a marketplace --- but also a forge, a campfire, a brotherhood of mud and steel. The Black Banners are not a single army but a squabbling league of freelance violence, bound by camp law, customs of compensation, and the shared knowledge that tomorrow's enemy might be today's employer. The real battles are fought in tents over agreements; the clash outside is just the review. Loyalty is a bond that shifts with circumstance --- but it also lives in the sergeant who shares his last ration, the surgeon who works through the night, the captain who stands in the shield wall.

Between the rigid honor codes of Ecktorian nobility and the fluid pragmatism of steppe clans and hill tribes, mercenaries navigate a dangerous middle ground. Survival depends on knowing when to keep an oath and when to break one --- always at dawn, with a witness, and with surety paid. But it also depends on knowing whose back to watch, whose flask to share, and whose name to call when the snow closes in.

\subsection*{Names of the Black Banners}
Mercenary names are often aliases, nicknames from past campaigns, or the color of a faded banner. Surnames are rare; epithets are everything.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Banner Names} & \textbf{Epithets} \\
\hline
Rurik & Svetlana & Blackwolf & \textit{the Unpaid} \\
\hline
Dragan & Mira & Ironhand & \textit{the Scarred} \\
\hline
Sebastian & Olena & Rustcloak & \textit{Bell-Heart} \\
\hline
Vladan & Zoya & Ashbanner & \textit{the Silent} \\
\hline
Temur & Nadezhda & Greyboar & \textit{Six-Fingers} \\
\hline
Caspian & Rada & Redtally & \textit{the Lame} \\
\hline
\end{longtable}

\subsection*{Quickstart (Two-Minute Setup)}
\label{sec:black-banners-quickstart}

\begin{enumerate}[leftmargin=*]
    \item \textbf{Condotta Crossroads:} Put two conflicting charters on the table. Ask: \textit{Who do you disappoint first?}
    \item \textbf{Company Sheet:} Name, Colors, Captain. Add timers: \textbf{Payday [4]}, \textbf{Supplies [4]}, \textbf{Cohesion [6]}.
    \item \textbf{Pressure Timers:} Pick two external threats (e.g., \textit{Skirmish Spiral [6]}, \textit{Camp Fever [6]}, \textit{Winter Deepens [6]}).
    \item \textbf{Allegiances in Flux:} At each scene start, roll 1d6 and apply the twist (see table).
    \item \textbf{Honor vs. Pragmatism Dial:} Slide based on choices; affects Position, logistics, and how the men remember you.
\end{enumerate}

\subsection*{The Four-Suit Generator}

Draw until you have all four suits. Highest rank sets the main Timer (2–5 $\rightarrow$ 4, 6–10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10). Any Ace echoes winter \& loyalty motifs: hunger, omen, promotion-by-attrition, or a sudden blizzard that buries both sides.

\subsubsection*{Spades --- Places (Camps, Battlefields, Fortresses)}

\begin{longtable}{|p{0.08\textwidth}|p{0.2\textwidth}|p{0.67\textwidth}|}
\hline
\textbf{Rank} & \textbf{Name} & \textbf{Description \& Mechanical Hook} \\
\hline
2 & Burned Village & Field hospital; flies, tinctures, fever-prayers. `[DEATH]` \textit{Start a \textbf{Camp Fever [4]} timer. Each failed Medicine roll ticks it. The dead whisper the names of the careless.} \\
\hline
3 & Frozen Ford & Last month's dead held under glass; standards trapped in the ice. `[DEATH]` `[WEATHER]` \textit{Crossing costs +1 Fatigue. The dead whisper a name --- the next \textbf{Desertion} check is at –1 die.} \\
\hline
4 & Latrine Row & Mud, shame, whispered mutiny; a sergeant who hears too well. `[SOCIAL]` \textit{Any failed social roll here ticks \textbf{Cohesion +1} (soldiers lose faith). The smell is a warning.} \\
\hline
5 & Crater Pasture & Alchemical barrage aftermath; glass flowers sprout like guilt. `[MAGIC]` \textit{Scavenging here gives a \textbf{Glass Shard} (1 use: +1 die to intimidate), but ticks \textbf{Supplies +1} (poisoned ground). Veterans avoid it.} \\
\hline
6 & Supply Road & Twelve broken wagons; oxen nose the wreckage for salt. `[TRAVEL]` \textit{First scavenge roll of the scene is \textbf{Controlled}; on a Miss, tick \textbf{Supplies +1}. The crows are fat.} \\
\hline
7 & Half-Raised Siege Tower & Abandoned camp; every rung creaks with grudges. `[COMBAT]` \textit{Climbing costs 2 Fatigue; on a Partial, the tower leans --- next attack from it is \textbf{Desperate}. The ghosts of sappers still work the ropes.} \\
\hline
8 & Salt-Cured Cache & Behind the surgeon's tent; wrong kind of maggots move. `[DEATH]` \textit{Opening it gives 1 \textbf{Supplies} but starts a \textbf{Disease [4]} timer (the maggots carry rot). The surgeon's face goes pale.} \\
\hline
9 & Command Tent & Maps show borders diplomacy erased. `[SOCIAL]` \textit{A player who studies the map for a scene may ask one question about enemy intentions; on a Miss, the map is a decoy --- start \textbf{False Orders [4]}. The tent smells of fear and bad wine.} \\
\hline
10 & Watch-Tower & Stacked from skulls and slate; crows know the countersign. `[WAR]` \textit{The crows will warn of ambush once, but only if fed a memory (erase a minor fact). The watchman never sleeps.} \\
\hline
J & Ancient March-Keep & Held by a skeleton garrison of veterans; they remember when banners were gold. `[DEATH]` \textit{They will fight for you if you swear an oath to restore their banner --- start \textbf{Oath Share [6]}. Their captain has no shadow.} \\
\hline
Q & The Bone Fields & Ykrul elders parley among cairns; reckon oaths with the dead. `[DEATH]` `[OATH]` \textit{The dead vote on any dispute. Roll Spirit+Lore (DV 4). On success, they side with you (gain +1 Position on next parley). On failure, they demand a blood-price. The wind carries their voices.} \\
\hline
K & Marcher's Fortress & Three walls, two sieges' scars, one gate that opens inward only. `[COMBAT]` \textit{The gate opens inward so the dead can leave. Living who enter without permission take Harm 1 (spiritual backlash). The stones weep when an agreement is broken.} \\
\hline
A & The Singing Wastes & Fallen weapons hum with voices of their last wielders. `[MAGIC]` `[DEATH]` \textit{Each weapon can answer one question about its owner's death. After three questions, the Wastes awaken --- start \textbf{Wrath of the Fallen [8]}. The song is a lament.} \\
\hline
\end{longtable}

\subsubsection*{Hearts --- People \& Factions}

\begin{longtable}{|p{0.08\textwidth}|p{0.2\textwidth}|p{0.67\textwidth}|}
\hline
\textbf{Rank} & \textbf{Name} & \textbf{Description \& Mechanical Hook} \\
\hline
2 & Young Cadet & Father's sword, mother's share, hunger to prove himself. `[SOCIAL]` \textit{If you mentor him (a downtime scene), he becomes a Cap 1 follower. If you ignore him, he joins a rival company --- tick \textbf{Rival Agenda +1}. He has never seen a man die.} \\
\hline
3 & Veteran Sergeant & Counts scars like medals; spends them on whiskey and mercy. `[COMBAT]` `[SOCIAL]` \textit{She will fight for you, but at end of scene demands a story about a battle you survived. If you cannot provide one, she takes a button from your coat --- a promise.} \\
\hline
4 & Ecktorian Banner-Captain & Griffon standard, roster of noble grudges, smile like winter. `[LAW]` \textit{He offers a \textbf{Letter of Marque} (once: ignore one legal complication). Accepting starts his \textbf{Favor Share [4]} --- he will call in a favor during a critical moment. His boots are always clean.} \\
\hline
5 & Condotta Broker & Changes colors when wind turns, always lands on his feet. `[SOCIAL]` \textit{He will flip a contract for you --- but the price is a \textbf{Black Mark} (reputation stain). He has three names and uses none.} \\
\hline
6 & Ykrul Clan Emissary & Speaks in riddled proverbs; pays in copper from a cursed mine. `[SOCIAL]` `[OATH]` \textit{He offers a \textbf{Safe-Conduct Token} (one free passage through Ykrul land). Accepting marks \textbf{Obligation [2]} --- the horse you ride will remember.} \\
\hline
7 & Vilikari War-Chief's Second Son & Honor burns brighter than his father's patience. `[SOCIAL]` \textit{He will join your next battle as a Cap 2 ally, but if he takes Harm, start a \textbf{Blood Feud [6]} with his clan. His axe has a name.} \\
\hline
8 & Surgeon of the Red Tally & Treats friend and foe; her record balances in blood. `[SHARE]` \textit{She can heal 1 Harm --- but the patient's name is added to her roll. Once per arc, she calls in the share: a dangerous mission. Her hands never shake.} \\
\hline
9 & Quartermaster & Knows where every coin went and every man will fall. `[SHARE]` \textit{His bribe book is thicker than his roster. Spend 1 Boon to cancel one \textbf{Supplies} tick --- but he now has a String on you. He never eats with the men.} \\
\hline
10 & Tribune's Aide-de-Camp & Sealed orders contradict the general's intent. `[SOCIAL]` \textit{He will leak the real orders for a \textbf{Favor Share [4]} --- but the leak may be a trap. His seal ring is warm.} \\
\hline
J & The Black Colonel & Legendary commander not seen in months, yet his banner rides. `[LEGEND]` \textit{The banner rides alone. If you follow it, you find a cache (gain 1 \textbf{Diamond} of your choice). Following it three times summons his ghost. He never lost a battle --- only his way home.} \\
\hline
Q & Ykrul Clan-Mother & Weaves war-braids from hair of oath-breakers. `[OATH]` \textit{She can witness any oath; breaking it gives her power over you (she may demand a service). Her braids have nine knots.} \\
\hline
K & Vilikari High Chief & Trades mercenaries like game pieces; never tips the board. `[SOCIAL]` \textit{He offers a \textbf{Trading Charter} (buy/sell anything in any camp). Using it ticks \textbf{Cohesion +1} (soldiers suspect you of corruption). His smile is an agreement.} \\
\hline
A & The Bannerless One & Commands loyalty without rank or flag. `[LEGEND]` \textit{She appears when the camp is darkest. She clears one segment of the pay cycle --- but you incur a \textbf{Price of Mercy [4]}. When it fills, she asks for a prisoner --- or a promise. Her face is never the same twice.} \\
\hline
\end{longtable}

\subsubsection*{Clubs --- Complications/Threats}

\begin{longtable}{|p{0.08\textwidth}|p{0.2\textwidth}|p{0.67\textwidth}|}
\hline
\textbf{Rank} & \textbf{Name} & \textbf{Description \& Mechanical Hook} \\
\hline
2 & Payday Delayed & Supplies dwindle before coin arrives. `[SHARE]` \textit{Tick \textbf{Payday +1}. On fill, a squad deserts --- lose 1 \textbf{Supplies} and gain a \textbf{Black Mark}. The captain blames the weather.} \\
\hline
3 & False Orders & Your target is now your ally --- on paper. `[SOCIAL]` \textit{The first command this scene is misinterpreted unless countered with a \textbf{Diamond} (cipher, cantor, witness). The messenger's hands are clean.} \\
\hline
4 & Blizzard Pins & Two rival companies, one cloister; vows and knives both come out. `[WEATHER]` `[COMBAT]` \textit{Tick \textbf{Cohesion +1}. The abbot will sell safe passage to the highest bidder --- start an \textbf{Auction [4]} timer. The snow is red by dawn.} \\
\hline
5 & Ykrul Counter-Offer & Better terms than Ecktoria; your contract becomes a liability. `[SOCIAL]` `[SHARE]` \textit{You may accept (gain +1 Effect, mark Obligation [2]) or refuse (tick the pay timer +1). The horses watch you decide.} \\
\hline
6 & Condotta Flip & Unit switches colors mid-battle; your ``relief'' hits your flank. `[COMBAT]` \textit{On a Miss, the unit attacks you. On a Partial, they demand double compensation --- tick \textbf{Payday +1}. Their old banner burns in your camp.} \\
\hline
7 & Camp Fever & Blooms in the surgeon's tent; the counting-house for the dead. `[DEATH]` \textit{Start \textbf{Camp Fever [4]}. Each failed Endurance test ticks it. At 4, a key NPC dies. The surgeon weeps in silence.} \\
\hline
8 & Vilikari Raid & Cuts your supply road; winter moves the schedule. `[COMBAT]` `[TRAVEL]` \textit{Tick \textbf{Supplies +1}. The raiders left a recipe --- cooking it gives 1 \textbf{Supplies} but starts a \textbf{Rival Hunt [4]} (they want the recipe back). The stew tastes of revenge.} \\
\hline
9 & Honor Code Collision & Ecktorian ideals vs. mercenary pragmatism; the camp splits by doctrine. `[SOCIAL]` \textit{Choose \textbf{Honor} (tick \textbf{Supplies +1}, logistics suffer) or \textbf{Pragmatism} (gain +1 Effect now, but start \textbf{Parley Divided [4]}). The men watch which way you lean.} \\
\hline
10 & Tribune Captured & The aide must choose: the rescue or the war. `[SOCIAL]` `[SHARE]` \textit{Rescue him: lose 1 \textbf{Supplies}. Abandon him: gain \textbf{Black Mark} but tick the pay timer –1 (his captors compensate you). His ghost will remember.} \\
\hline
J & Black Colonel's Banner & Appears on the wrong field; confusion is a weapon. `[COMBAT]` \textit{Roll 1d6. On 1–3, the enemy believes you have reinforcements (+1 Position). On 4–6, your own side hesitates (–1 die on first attack). The banner has no shadow.} \\
\hline
Q & Blood-Feast & Clan-Mother calls it; all contracts suspended until dawn's judgment. `[OATH]` `[SOCIAL]` \textit{The feast is a trial. Each PC must confess a broken oath or take \textbf{Harm 1} (the clan judges). The meat is always fresh.} \\
\hline
K & High Chief's Game & Plays companies against each other; ``winner'' takes obligations, not spoils. `[SOCIAL]` `[SHARE]` \textit{The ``winner'' inherits a \textbf{Obligation Timer [6]} from the loser. Refusing the game starts a \textbf{Vendetta [4]}. The game has no rules --- only winners.} \\
\hline
A & The Singing Wastes Awaken & Every fallen blade rises to seek its last hand. `[MAGIC]` `[DEATH]` \textit{The blades attack everyone for 1 round. Afterward, any PC may claim a \textbf{Cursed Blade} (+1 die, but the blade whispers secrets --- the GM gains 1 SB per scene). The song is a scream.} \\
\hline
\end{longtable}

\subsubsection*{Diamonds --- Treasures \& Discoveries}

Each Diamond is a physical object with a cost and expiration.

\begin{longtable}{|p{0.08\textwidth}|p{0.2\textwidth}|p{0.67\textwidth}|}
\hline
\textbf{Rank} & \textbf{Name} & \textbf{Description \& Mechanical Hook} \\
\hline
2 & Emergency Requisition & Strip a fallen company; their ghosts watch. `[SHARE]` \textit{Gain 1 Use of any mundane gear. After use, the quartermaster's ghost haunts you for 1d4 scenes (–1 die on stealth). He asks for his boots back.} \\
\hline
3 & Dual Contract & Serve two masters whose goals cannot both be met. `[SOCIAL]` `[SHARE]` \textit{Gain +2 dice on one negotiation, but start a \textbf{Double Charter [6]} --- when it fills, both masters seek recompense. The second master always cheats.} \\
\hline
4 & Honor Dispensation & Your acts are justified regardless of code (for now). `[SOCIAL]` \textit{Once, ignore a \textbf{Black Mark} consequence. The dispensation is a bell --- ring it, and the tribunal looks away. The bell then cracks. The sound is a lie.} \\
\hline
5 & Condotta Rider's Pass & Cross any camp perimeter unchallenged. `[STEALTH]` \textit{The pass is a feather from a messenger who arrived too late. Use it once; then the feather burns. The smoke spells a name.} \\
\hline
6 & Ykrul Safe-Conduct Token & Pass a named clan's lands unmolested (until the moon changes). `[TRAVEL]` \textit{The token is a horse's tooth. The horse was a gift from a dead Khagan. The tooth turns to dust after one use. The horse still whinnies at midnight.} \\
\hline
7 & Vilikari War-Mark & Their scouts guide you through hostile hills --- and remember the favor. `[TRAVEL]` `[SHARE]` \textit{Gain +1 Effect on one navigation roll. The favor will be called at dawn: they will demand a ridge or a bridge. The ridge is haunted.} \\
\hline
8 & Surgeon's Note & Medical care when no coin remains; the interest is names. `[SHARE]` \textit{Heal 1 Harm now. The note is a name --- the surgeon will collect it later (GM gains a String on you). She never forgets a face.} \\
\hline
9 & Tribune's Cipher Ring & Decode orders that do not exist yet. `[SOCIAL]` \textit{Once, treat a \textbf{False Orders} complication as \textbf{Clean Success}. The ring is warm --- it cools after use. The message was always a lie.} \\
\hline
10 & Banner-Captain's Seal & Commandeer resources from junior companies (and their resentment). `[AUTHORITY]` \textit{Gain +1 \textbf{Supplies} now, but start a \textbf{Jealousy [4]} timer (junior officers scheme against you). The seal is a wolf's head. It bites.} \\
\hline
J & Field Promotion Warrant & Assume command of a shattered unit, inherit its timers. `[AUTHORITY]` `[SHARE]` \textit{You gain a Cap 2 follower (the unit's sergeant), but inherit their pay timer [2/4] and \textbf{Black Mark [1]}. The sergeant has seen too much.} \\
\hline
Q & Clan-Mother's War-Braids & Her chosen fight beside you this battle, then judge you after. `[COMBAT]` `[SOCIAL]` \textit{Gain a Cap 3 ally for one battle. Afterward, she judges your conduct --- on a Miss, she takes a war-braid from you (lose 1 Boon). The braid is warm.} \\
\hline
K & High Chief's Trading Charter & Buy or sell \emph{anything} in any camp, no questions asked aloud. `[TRADE]` \textit{Once per arc, treat a market roll as \textbf{Clean Success}. The questions are asked in silence --- the answers are compensated in silence (GM gains 1 SB). The charter is written on skin.} \\
\hline
A & The Bannerless Word & Loyalty that outruns rank, law, and coin. `[LEGEND]` \textit{Once, when all other resources are exhausted, whisper this word to rally a shattered company. They fight for one scene without tracking the pay timer or cohesion. Afterward, the word is forgotten --- you cannot use it again. The men will not speak of it.} \\
\hline
\end{longtable}

\subsection*{Company Mechanics (Visible Tracks)}

Every mercenary company (PC-led or NPC) tracks three timers:

\begin{itemize}
    \item \textbf{Payday [4]} – Missed compensation, bad terms, seized chests. On fill: a squad deserts (lose 1 \textbf{Supplies}) or a veteran sells secrets (GM gains 2 SB). \textit{The men whisper.}
    \item \textbf{Supplies [4]} – Weather, marches, sabotage. While Supplies $\geq$ 3, all risky maneuvers start \textbf{Position –1} unless you bring a Diamond (scavenge, requisition). \textit{Hunger is a slow knife.}
    \item \textbf{Cohesion [6]} – Losses, betrayals, officer feuds. When Cohesion hits 1–2, the GM may convert the first 1 SB each scene into \textbf{Desertion, Brawl, or Refusal}. \textit{Trust is the first casualty.}
\end{itemize}

\textbf{Recovery:} A successful \textbf{Payline Drill} scene clears 1 Payday. A \textbf{Quartermaster Audit} clears 1 Supplies. A \textbf{Rally \& Reform} roll (Presence+Command, DV 3) clears 1 Cohesion. \textit{Shared hardship binds faster than coin.}

\subsection*{Camp Event Mini-Generator (d8)}
\label{sec:black-banners-camp-events}

When the company makes camp or when tensions run high, roll on this table for a camp event that affects morale, cohesion, or supplies.

\begin{longtable}{|p{0.1\textwidth}|p{0.85\textwidth}|}
\hline
\textbf{d8} & \textbf{Event} \\
\hline
1 & \textbf{Deserter's Tale.} A soldier tries to flee with a payroll satchel. If caught, the company demands a public flogging. Intervening costs 1 Boon but earns a loyalty share (+1 Cohesion). If ignored, tick \textbf{Payday +1} (the coin is gone) and \textbf{Cohesion –1} (trust erodes). The deserter's name is scratched from the roll. \\
\hline
2 & \textbf{Surgeon's Lottery.} The camp surgeon runs a lottery for medical priority. Those who pay 1 Boon jump the queue. Those who refuse may lose a wounded ally to infection (lose 1 \textbf{Supplies} or a follower). The surgeon's ledger grows longer --- and her smile thinner. \\
\hline
3 & \textbf{Muster of the Fallen.} A ghostly formation marches through camp at midnight. Veterans take it as an omen. Each PC must test Resolve (DV 3) or gain the \textbf{Haunted} condition (–1 die on initiative for one day). On a success, they gain +1 die on the next morale check. The dead are counting. \\
\hline
4 & \textbf{Brawl Over Bread.} A fight erupts over the last rations. The quartermaster demands the party intervene. Choose: \textit{Break it up} (costs 1 Fatigue, no change), \textit{Take sides} (gain a small faction's favor but tick \textbf{Cohesion –1}), or \textit{Let it burn} (lose 1 \textbf{Supplies} to spoilage). The loser's face stays in the mud. \\
\hline
5 & \textbf{Pardoner's Visit.} A traveling pardoner offers absolutions for past sins. Pay 1 Boon to clear one \textbf{Black Mark} (the pardoner writes your name in his book). Refuse, and the pardoner's curse gives –1 Position on the next social roll against clergy. His book has many names. \\
\hline
6 & \textbf{Desertion Ring.} An informant reveals a plot to desert en masse. The party can pay arrears (clear 1 Payday), execute the ringleaders (gain a \textbf{Black Mark} but tick \textbf{Cohesion +1}), or ignore it (tick \textbf{Payday +2} as half the camp vanishes overnight). The informant wants a horse. \\
\hline
7 & \textbf{Camp-Fever Peak.} The surgeon's tent overflows. Mark 1 \textbf{Fatigue} for each PC. A successful Medicine test (DV 4) reduces the effect by half. On a critical failure, a key NPC (the quartermaster or a sergeant) dies. The fever has a name. \\
\hline
8 & \textbf{The Bannerless One Visits.} The mysterious figure appears at the edge of the firelight. She offers to clear 2 pay cycle segments in exchange for a \textbf{Price of Mercy [4]} timer. The timer advances each time the party shows mercy to an enemy. When full, she asks for a prisoner --- or a promise. Her eyes are old. \\
\hline
\end{longtable}

\subsection*{Allegiances in Flux (Scene Start, 1d6)}

At the start of every scene involving mercenaries, roll 1d6:

\begin{enumerate}[leftmargin=*]
    \item \textbf{Condotta Flip} – Rival charter offers better terms. Accept: gain \textbf{Black Mark [1]} with old patron. Refuse: tick \textbf{Payday +1}. \textit{The ink is still wet.}
    \item \textbf{Ykrul Counter-offer} – Accept an exit-rich but costly route: gain +1 Effect now, mark \textbf{Obligation [2]}. Refuse: tick \textbf{Payday +1}. \textit{The wind shifts.}
    \item \textbf{Vilikari Shortcut} – +1 Effect on this maneuver via night roads; GM gains 1 SB (hostile terrain later). \textit{The path is older than the kingdom.}
    \item \textbf{Ecktorian ``Correction''} – March under ``imperial guidance'': +1 Position this scene or take a \textbf{Black Mark} for defiance. \textit{The seal is heavy.}
    \item \textbf{False Orders} – First command this scene suffers misinterpretation unless countered with a \textbf{Diamond} (cipher, cantor, witness). \textit{The messenger's hands are clean.}
    \item \textbf{Bannerless One} – A myth-wrapped figure clears 1 pay segment – but starts \textbf{Price of Mercy [2]}. When it fills, she asks for a prisoner you'd rather keep. \textit{She was always there.}
\end{enumerate}

\subsection*{Honor vs. Pragmatism Dial}

Place a two-step dial. Starting position: balanced (0).

\begin{itemize}
    \item \textbf{Honor} (left): +1 Position in parley/oathbound actions; logistics burn faster (increase \textbf{Supplies} by 1 on a Partial where you'd normally be safe). \textit{The old songs approve.}
    \item \textbf{Pragmatism} (right): +1 Effect in ambush/sabotage; social rolls start one step worse Position when oaths/witnesses matter. \textit{The men learn to watch their backs.}
\end{itemize}

\textbf{Centering Actions:} Settle arrears on-screen, return prisoners with witness, fulfill a broken writ at cost. Move the dial one step toward center. \textit{The campfire remembers.}

\subsection*{Black Banners Superstitions (burn for +1 die once)}
\begin{itemize}
    \item \textbf{Bread to the Bell:} Crumble a crust at a camp's bell post. Ignore the first \emph{Payday Delayed} penalty. [SUPERSTITION]
    \item \textbf{Knot the Roster:} Tie a single knot in the company roster. Cancel \emph{Condotta Flip} or take –1 die on your next negotiation (your choice). [SUPERSTITION]
    \item \textbf{Name to the Mud:} Whisper a dead captain's name into the latrine trench. Gain +1 Effect when rallying the company this scene, but mark \emph{Obligation} to the Bannerless One. [SUPERSTITION]
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
    \item \textbf{Advance/Withdraw (Wits + Ride or Stealth):} On a hit, shift one band. On a strong hit, also impose \emph{Dust Cloud} (cover retreat).
    \item \textbf{Cut the Contract (Presence + Sway):} On a hit, freeze range for one exchange; if a \emph{Condotta Rider's Pass} is in your possession, +1 die.
    \item \textbf{Weather the Blizzard (Resolve + Survival):} On a hit in \emph{Blizzard Pins}, you pick who finds shelter; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Three-Phase Foes)}
\begin{itemize}
    \item \textbf{The Coin-Captain (Mercenary Turncoat)} --- \emph{Phases}: Fair Offer $\rightarrow$ Betrayal Clause $\rightarrow$ Payday Massacre. \emph{Cracks}: expose his dual contract before the company, pay off his debt to the Condotta Broker, or outbid him for his own loyalty.
    \item \textbf{The Quartermaster-Provost (Logistics Tyrant)} --- \emph{Phases}: Supply Freeze $\rightarrow$ Ration Cuts $\rightarrow$ Starvation Siege. \emph{Cracks}: steal his secret ledger, rally the men with a shared meal (spend 1 Supplies), or break his control of the supply route.
    \item \textbf{The Singing Wakes (Weapon-Host)} --- \emph{Phases}: Whisper $\rightarrow$ Chorus $\rightarrow$ Scream. \emph{Cracks}: perform the \textit{Three Silencing Acts} (funeral honors for a fallen blade, bury a broken standard, ring the wind-break bell).
\end{itemize}

\subsection*{GM Tools}

\paragraph{Field Procedures}
\begin{itemize}
    \item \textbf{Colors \& Countersigns:} Declaring correct countersigns cancels the first \textbf{False Orders} SB each night; wrong signs worsen Position on first contact.
    \item \textbf{Payline Drill:} If a scene depicts orderly compensation, clear 1 \textbf{Payday}. If pay is ``creative,'' clear it but mark \textbf{Black Mark} – future officers begin one step colder in parley.
    \item \textbf{Scavenge Doctrine:} After battles, gain 1 temporary Gear tag or advance \textbf{Supplies} by 1; on a 1–2 on 1d6, spawn ``Honor vs. Pragmatism'' conflict immediately.
\end{itemize}

\subsection*{Cross-Regional Use}
\begin{itemize}
    \item \textbf{Vhasia:} A \emph{Condotta Rider's Pass} substitutes for a \emph{Safe-Conduct} on the Pont-du-Tithe once; after use, start a \emph{Parlement Divided} tick.
    \item \textbf{Viterra:} A \emph{Ykrul Safe-Conduct Token} may be notarised as \emph{Ferry Priority}; the notary's fee is a story.
    \item \textbf{Ubral:} A \emph{Clan-Mother's War-Braid} counts as a \emph{Guest-Token} in hill-forts; the clans will ask whose side you took.
    \item \textbf{Mistlands:} A \emph{Banner-Captain's Seal} can be presented to a Bell-Warden as proof of lawful passage; the Warden tests it with a bell. If it rings false, the mist takes a day.
    \item \textbf{Valewood:} A \emph{Field Promotion Warrant} can be offered to a Green Knight as proof of worth; the Knight may challenge you to a trial of paths instead of blades.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In the Black Banners, power flows through agreements, coin chests, and the weight of a reputation --- but also through the sergeant who knows each man's name, the cook who stretches the rations, the smith who works through the night. A Banner-Captain's authority rests on the number of companies he commands; a Condotta Broker's on the thickness of his roster. Yet loyalty is not bought --- it is earned in the mud, at the latrine, across a shared fire.

\textbf{For the GM:}  
Power in the Black Banners is measured in payday timers, broker's favors, and the length of a duel record --- but also in the camp's mood, the quartermaster's ledger, and the stories the veterans tell. Patronage often takes the form of dual contracts, emergency requisitions, or the Bannerless Word – each tied to a specific company, broker, or share. To emphasize the mercenary realism:
\begin{itemize}
    \item Use the pay and supply timers as visible pressure. Let players negotiate, raid, or pledge future services to clear ticks.
    \item Enforce \textbf{Allegiances in Flux} – every scene, loyalties may shift. The Keeper's SB pool grows when the party breaks an agreement.
    \item Let \textbf{Honor vs. Pragmatism} be a dial the party can slide through their actions. Center it by a costly act of mercy or a profitable betrayal.
    \item Track \textbf{Black Marks} as a persistent reputation score. Three Black Marks and no condotta broker will write a clear agreement for you.
    \item Remember the human element: the soldier who shares his last ration, the captain who stands in the line, the surgeon who weeps in private. These are the ties that outlast coin.
\end{itemize}
In the Black Banners, the camp sings when pay arrives --- and falls silent when the coin chest is empty. But the camp also sings when a wounded friend survives, when the supply train arrives, when the snow melts and the road opens. The song is never just about coin.
\end{tcolorbox}

\subsection*{Final GM Mantra (Black Banners Edition)}
The camp sings when pay arrives --- and falls silent when the chest is empty. Set \textbf{DV low}. Use \textbf{Allegiances in Flux} every scene. Let \textbf{Payday, Supplies, and Cohesion} be visible. Spend \textbf{SB} as false orders, cut supply lines, or camp fever. And remember: in the Black Banners, a broken agreement costs more than a broken blade --- but a broken trust costs more than either.

% Index entries
\index{Black Banners|see{Condotta \& Crowns}}
\index{Theme generator}
\index{Condotta \& Crowns generator}
\index{Places!Black Banners}
\index{People!Black Banners}
\index{Complications!Black Banners}
\index{Rewards!Black Banners}
\index{Banner-Captain}
\index{Condotta Broker}
\index{Quartermaster}
\index{Payday timer}
\index{Cohesion timer}
\index{Camp event generator}
\index{Allegiances in Flux}
\index{Honor vs. Pragmatism dial}
\index{Local Patrons!Black Banners}