% ============================================================
% Region Chapter Content
% ============================================================
\section{The Brass Coast --- ``The Shore of Silver and Spice''}
\label{chap:brasscoast}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{Three powers, one coast, and a memory written in salt. The wrong spark will burn the whole shore.}

\begin{quote}
  \textbf{Raja Govind (Native Hill Lord):}  
  ``My grandfather watched the Kahfagians raise their first customs house on our beach. My father paid their tolls in sweat. I will see them drown in their own silence — or I will drown trying. The Sidhi are guests. The Admiralty are thieves. The Brass Coast belongs to those who salt their own fish.''
\end{quote}

\begin{quote}
  \textbf{Sidhi Freedman (Commoner):}  
  ``My grandmother was born in Galanina. My mother was born in a slave ship. I was born in a lighthouse. The Admiralty calls me a subject. The Covenant calls me a freedman. But the rajas — they call me \textit{guest}. And a guest who stays too long becomes a target.''
\end{quote}

\textbf{Genre / Mood:} Mercantile noir, post-colonial pressure cooker, ghost-lit coast, three-way cold war.

\textbf{Starting Location:} A moonlit beach where three boats meet: a Kahfagian cutter, a Sidhi lighthouse skiff, and a Dhaharan spice dhow. A native pearl diver approaches you: ``The Admiralty raised the salt tax again. The Sidhi say they will strike the ropeways. And my raja is sharpening his father's sword. Pick a side — or the tide will pick for you.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Brass Coast is Dhahara's southern rim: a humid, feverish coastline of palm groves, salt pans, and crowded port cities. Three powers bleed into each other here, and none can draw a clean breath.

\noindent
\textbf{The Admiralty (Kahfagian)} holds the warm-weather ports and the toll schedules. Their warships patrol the shipping lanes, and their magistrates judge by maritime law. They call it a ``client province.'' The locals call it a noose. The Admiralty is not evil — it is \emph{hungry}. Its thalassocratic empire needs new markets, raw materials, and a captive shoreline to fuel its oligarchs' ever-growing appetites. The Brass Coast is a vein they intend to mine dry.

\noindent
\textbf{The Sidhi} are a foreign diaspora — refugees from Galanina who fled the Ashaani slavers generations ago. They control the ropeways, the lighthouse courts, and the freedman's seal. They are welcome nowhere and trusted less, but without them the spice trade stops. Their story is one of forced migration: driven from their homeland, they moved from place to place, unwelcome at every stop, until at last they reached this coast. Like the displaced of another age, they made a new life from salt water and stubborn memory. They resent the rajas for the cold shoulders and the whispers — ``foreigners,'' ``drifters,'' ``they keep to themselves.'' And they resent the Admiralty for trying to subjugate them all over again, trading chains of iron for chains of tolls. The Sidhi make the spice flow — and wherever flow is strong, hands reach to dam it.

\noindent
\textbf{The Native Dhaharan Rajas} are the hill lords who have lived here since before the Admiralty's first anchor. They grow the pepper, mine the iron, and salt the fish. They watch the coast from inland fortresses, and they remember when no foreign flag flew over their beaches. Some accept bribes. Some sharpen blades. All of them are waiting. They resent the Admiralty for the slow colonization — the tolls, the land seizures, the quiet erasure of their courts. But they also resent the Sidhi. Centuries have passed, yet the Sidhi remain ``outsiders'' in the rajas' eyes — a people who keep to their own lighthouses, marry within their own circles, and speak a creole tongue that grates on the ear. The rajas envy the Sidhi's quiet prosperity: the ropeways humming with trade, the seal that opens every port. They do not trust a people who have no homeland and yet have thrived. And that envy sharpens their knives.

The air smells of pepper, cinnamon, and the low tide of a thousand ships. Here, a man can buy anything — a ship, a spice lot, a freedman's seal — and sell anything except his own shadow. But the price is never just silver. The Admiralty takes tolls. The Sidhi take stories. The rajas take blood.

\vspace{2mm}
\textbf{What you notice first:} The heat pressing down like a damp cloth. The way the sea is the same color as the sky, so the horizon disappears. The brass hooks on Sidhi belts, the brass scales on Admiralty desks, and the brass earrings in native chieftains' ears — three metals, one coast.

\textbf{The one rule that kills newcomers:} Never board a ship that has a green lantern at the mast. It means the captain is under a death vow — and you are the payment. Also, never refuse a cup of palm wine offered by a raja's envoy. It is a test of honor, and the hills have long memories.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Toll, The Heat of the Hills}

The Brass Coast is not a colony — it is a slow-burning argument. The Admiralty holds the ports, the Sidhi hold the ropeways, and the rajas hold the land. Each side believes the other two are temporary. Each side is wrong.

\noindent
\textbf{The Admiralty's Powder Keg:} The Kahfagian governor, Admiral Voss, rules from a whitewashed fort that was once a Sidhi stronghold. His marines drink coconut wine and mock the ``lazy natives.'' His clerks adjust tolls with the tide. But his ships are few, his garrison is underpaid, and the rajas have started hiring pirate dhows to intercept his grain convoys. The Admiralty needs new markets because its old ones are failing; the Brass Coast is meant to be the solution. Instead, it has become a wound that will not close.

\noindent
\textbf{The Sidhi's Fragile Web:} The Sidhi run the ropeways, the lighthouse courts, and the freedman's seal. They are the only neutral party — and neutrality is a knife edge. A Sidhi strike would collapse the spice trade. A Sidhi betrayal would hand the coast to the Admiralty. And the rajas whisper that the Sidhi are just another foreign master, trading one chain for another. The Sidhi do not forget the long road that brought them here — the forced marches, the closed gates, the children buried in unnamed graves. They will not be pushed again. But the rajas' envy and the Admiralty's hunger leave them few roads out.

\noindent
\textbf{The Rajas' Slow Burn:} The native Dhaharan hill lords speak a dozen dialects and worship a hundred spirits. They do not agree on much — except that the coast belongs to them. Some, like Raja Govind of the Kalinga foothills, play the Admiralty and the Sidhi against each other. Others, like the Veiled Queen of the Mangrove Delta, have already sent envoys to the Ashaani Breath-less. And in the smoky kitchens of every hill fort, young warriors sharpen their grandmother's steel and wait for the sign. They resent the Admiralty's tolls, yes — but they also resent the Sidhi's quiet prosperity. ``They do not marry our daughters,'' a raja's son might say. ``They do not pray to our gods. They are strangers who have grown rich on our land.'' That resentment is a fire the Admiralty fans when it suits them.

\noindent
The green flame of the Sidhi lighthouses, the brass scales of the Admiralty, and the iron swords of the rajas — the Brass Coast is a tripwire. Any spark could set it ablaze.

\subsection*{The Three Powers: Faction Clocks}
Use these timers to track the coast's simmering tensions.

\begin{itemize}
\item \textbf{Admiralty Grip [6]:} Ticks up when the party enforces Kahfagian tolls, sinks a pirate dhow, or exposes a Sidhi smuggler. At 4: Admiral Voss offers the party a Letter of Marque (legal piracy against rival powers). At 6: He declares martial law — all ports closed, all ships searched, all outsiders suspect.
\item \textbf{Sidhi Ropeway Control [6]:} Ticks up when the party helps a Sidhi lighthouse keeper, frees a Sidhi prisoner, or protects a Sidhi caravan. At 4: The Sidhi offer a Lighthouse Charter (safe harbour in any Sidhi port). At 6: They call a ropeway strike — trade halts, and every faction blames each other.
\item \textbf{Raja's Rebellion [8]:} Ticks up when the party smuggles weapons to the hills, assassinates an Admiralty official, or reveals a Sidhi secret. At 4: A raja sends a secret envoy with a tiger's claw (an invitation to an alliance). At 8: The hills rise — native war parties raid coastal towns, and the Admiralty responds with fire ships.
\end{itemize}

These timers are not mutually exclusive. A clever party might feed one to delay another — or light all three at once.

\subsection*{Prominent NPCs of the Brass Coast}
\label{sec:brasscoast-npcs}

\begin{longtable}{|p{0.22\textwidth}|p{0.28\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Admiral Voss & Kahfagian Governor & Extract every possible toll for the Admiralty & He keeps a record in his own blood; he says ink fades. \\
\hline
Syndic Olena & Port-master of Tamralipta (Sidhi) & Enforce the toll schedule — but protect her people & Her brass hook is also a stamp; the stamp leaves a scar. \\
\hline
Raja Govind & Native hill lord of the Kalinga foothills & Break the Admiralty's monopoly — and the Sidhi's influence & He has never seen the sea. He is afraid of it, but he will never admit that. \\
\hline
The Veiled Queen & Native ruler of the mangrove delta (matriarch) & Keep her people free by playing all sides against each other & She wears a mask of hammered silver; no one has seen her face in twenty years. \\
\hline
Captain Idris & Sidhi lighthouse commander & Keep the Brass Coast free of Ashaani slavers — and his people alive & His left eye is a brass lens; he can read the stars even in fog. \\
\hline
Shipwright Dara & Builder of freedmen's boats (Sidhi) & Launch one new vessel each year for escaped slaves & She names each boat after a dead family member. The names are never painted — only carved. \\
\hline
The Black Bishop & Covenant Maritime Judge & Enforce the law — but his conscience is a rusty nail & He wears a ring made from a freed slaver's ship nail. The nail still has rust on it. \\
\hline
Rani Kumar (Young Firebrand) & Raja Govind's daughter, leader of the hill youth & Drive the foreigners into the sea — by any means & She has a tattoo of a coiled tiger on her right arm. It moves when she is angry. \\
\hline
\end{longtable}

\subsection*{Names of the Brass Coast}
\label{sec:brasscoast-names}

Native Dhaharan names dominate (regional dialects). Sidhi names (Levantine/coastal) are marked with an asterisk. Kahfagian names are rare among locals.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surname/Clan} & \textbf{Epithets} \\
\hline
Govind & Olena* & Pallava & \textit{the Brass-Fingered} \\
\hline
Kannan & Dara* & Chettiar & \textit{Coin-Counter} \\
\hline
Muthu & Mangai & Saliya & \textit{the Tide-Reader} \\
\hline
Rajendran & Ilavarasi & Pattinam & \textit{Lighthouse Born*} \\
\hline
Senthil & Kalyani & Mudaliar & \textit{Shark-Tooth} \\
\hline
Anjali* & Kumar (surname) & Vellalar & \textit{the Unbent} \\
\hline
\end{longtable}

\paragraph*{(Harbour / Hill Fort / Lighthouse)} Tamralipta harbour (a forest of masts); Raja Govind's hill fort (laterite stone, painted with tigers); Sidhi lighthouse with a brass mirror and green flame.

\section*{Spades — Places}
\label{sec:brasscoast-places}

\noindent
\textit{Native sites are marked with \textdagger; Sidhi sites with *; Admiralty sites with \textdaggerdbl.}

\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Tamralipta Harbour*\textdaggerdbl} --- A forest of masts; the smell of tar and spices.
  \hfill \textit{The harbour is a maze of floating houses. The residents are all Sidhi. They never touch the shore — but they pay tolls to the Admiralty.} \texttt{[TRADE]}
\item \textbf{Pepper Warehouse (Native-owned)\textdagger} --- A long building with slatted walls; the air is sharp.
  \hfill \textit{The workers are freedmen earning passage for their families. Their eyes are red from pepper dust. They never blink. The owner is a raja's cousin.} \texttt{[LABOR]}
\item \textbf{Lighthouse of the Brass Flame*} --- A stone tower with a copper mirror.
  \hfill \textit{The light is green. It shows the faces of the drowned. The keeper knows their names — and some of them are still walking.} \texttt{[JUSTICE]}
\item \textbf{Salt Pan of the Silent Widows\textdagger} --- A flat expanse of white crust; women in black work the rakes.
  \hfill \textit{The women are all widows; their husbands were lost at sea — to storms, to slavers, to Admiralty press gangs. The salt is bitter. They sell it to remember.} \texttt{[COMMUNITY]}
\item \textbf{Kahfagian Admiralty House\textdaggerdbl} --- A white building with a veranda; a flag with a lantern.
  \hfill \textit{The flag is lowered at sunset. The Admiral is drunk by dawn. The clerks are sober — and watching. Native servants spit in the coffee.} \texttt{[AUTHORITY]}
\item \textbf{Sidhi Court of the Lighthouse*} --- A circular chamber with a brass scale.
  \hfill \textit{The judge is blind; she hears truth in the echo of a coin. The coin is never returned. The rajas send observers, but they never speak.} \texttt{[JUSTICE]}
\item \textbf{Palm Grove of the Elephants\textdagger} --- A grove where temple elephants are bathed.
  \hfill \textit{The elephants are painted with turmeric. The paint is a prayer for safe passage — and for the foreigners to leave.} \texttt{[RITUAL]}
\item \textbf{Coral Reef of the Sunken Ship (Neutral)} --- A reef just offshore; a mast breaks the surface.
  \hfill \textit{The ship was a slaver; the slaves escaped. They live in the reef now. They trade salvage for silence — and they remember which raja helped them.} \texttt{[LEGEND]}
\item \textbf{Pepper Auction Floor (Native-owned)\textdagger} --- A crowded hall; voices bid in claps.
  \hfill \textit{The auctioneer has a hammer of whalebone. It cracks when he lies. The cracks are getting louder. The Admiralty's toll agents sit in a raised gallery.} \texttt{[TRADE]}
\item[J] \textbf{The Green Lantern Tavern (Sidhi)*} --- A low building with a green lantern.
  \hfill \textit{The lantern is never lit. The owner says it would attract the wrong kind of dead. The right kind drink here for free — and they remember every whispered plot.} \texttt{[SOCIAL]}
\item[Q] \textbf{The Shipbreaker's Yard (Admiralty-licensed)\textdaggerdbl} --- A beach where old ships are dismantled.
  \hfill \textit{The workers are convicts; they wear chains that drag in the sand. They earn one link off each season. The chains are paid for by the rajas' taxes.} \texttt{[LABOR]}
\item[K] \textbf{The Admiral's Palace (Former Sidhi Fort)\textdaggerdbl} --- A former Sidhi fort, now occupied by Kahfagian officers.
  \hfill \textit{The flag is flown upside down. The Admiral does not know the difference. The Sidhi do. They count the days. The rajas count the cannons.} \texttt{[AUTHORITY]}
\item[A] \textbf{The Undercliff (Smuggler's Cove, native-controlled)\textdagger} --- A smuggler's cove accessible only at low tide.
  \hfill \textit{The entrance is a cave; the cave leads to a hidden harbour. The harbour is a city of thieves — and a union of freedmen. The rajas' tax collectors meet here to plan.} \texttt{[LEGEND]}
\end{enumerate}

\section*{Heads — People \& Factions}
\label{sec:brasscoast-people}

\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Rani Kumar (Native Firebrand)\textdagger} --- Raja Govind's daughter, with a tiger coiled on her arm.
  \hfill \textit{``My father negotiates. I sharpen swords. The Admiralty will leave — or they will be buried here. There is no third way.''} \texttt{[POLITICS]} \texttt{[REBELLION]}
\item \textbf{Sidhi Lighthouse Keeper*} --- Blind in one eye; carries a brass telescope. He can see the future through the lens.
  \hfill \textit{``Only the futures that end at sea,'' he says. ``The rest are private. I sell them to captains who need to avoid them — and to rajas who need to know when to strike.''} \texttt{[JUSTICE]}
\item \textbf{Kahfagian Port-Master (Syndic Olena)\textdaggerdbl} --- A thin woman with a brass hook for a hand. The hook is a stamp.
  \hfill \textit{She stamps every toll receipt. The stamp leaves a scar. She says it builds character. The scarred merchants say it builds grudges.} \texttt{[AUTHORITY]}
\item \textbf{Pepper Factor (Native)\textdagger} --- A thin woman with red-stained fingers and a black tongue. She is the raja's eyes in the market.
  \hfill \textit{She can tell the quality of pepper by taste. She has never been wrong. She has also never been paid on time — by the Admiralty.} \texttt{[TRADE]} \texttt{[SPY]}
\item \textbf{Shipwright Dara (Sidhi)*} --- A woman with a hammer and a jar of brass nails. She builds boats for freed slaves.
  \hfill \textit{Each boat is named after a dead family member. The names are never painted — only carved. The wood remembers. So do the rajas, who buy her boats for their own smugglers.} \texttt{[COMMUNITY]}
\item \textbf{Admiral Voss\textdaggerdbl} --- A grey-haired man with a glass eye. The glass eye is from a defeated pirate.
  \hfill \textit{The pirate still watches through it. Voss does not sleep well. The eye weeps salt water. He dreams of the rajas' torches.} \texttt{[AUTHORITY]}
\item \textbf{Raja Govind (Inland Hill Lord)\textdagger} --- A young man in silk, come to the coast to sell iron. He has never seen the sea.
  \hfill \textit{He is afraid of it. He refuses to step on a boat. He negotiates from the shore, shouting through a brass horn. His daughter mocks him — but she will not leave his side.} \texttt{[POLITICS]} \texttt{[TRAGEDY]}
\item \textbf{The Veiled Queen (Mangrove Delta)\textdagger} --- A woman in a silver mask, ruler of the southern marshes.
  \hfill \textit{``I have spoken to the Admiralty's spy, the Sidhi's whisperer, and the Ashaani's shadow. All of them will pay for my people's freedom — or I will drown them all in the delta. The mask is not to hide my face. It is to hide my smile.''} \texttt{[MYSTERY]} \texttt{[MASTERMIND]}
\item \textbf{Smuggler Captain (Neutral, works for highest bidder)} --- A woman with a parrot on her shoulder, a curved knife on her belt.
  \hfill \textit{The parrot is her navigator; it knows the reefs by sound. The parrot is blind. It has never been wrong. Her loyalty is to the cargo's weight, not any flag.} \texttt{[STEALTH]}
\item[J] \textbf{The Black Bishop (Covenant Maritime Judge)} --- A judge who rules on maritime disputes. He wears a ring made from a ship's nail.
  \hfill \textit{The ship was the first slaver he freed. The nail still has rust on it. He calls it his conscience. The Admiralty calls him a fool. The rajas call him a useful idiot.} \texttt{[JUSTICE]}
\item[Q] \textbf{The Green Lantern's Owner (Sidhi widow)*} --- A woman with no shadow. She sold her shadow for the tavern.
  \hfill \textit{The shadow now serves drinks. The customers do not notice. The shadow remembers every secret — and whispers them to the highest bidder.} \texttt{[MYSTERY]}
\item[K] \textbf{The Sea King of Kalinga (Pirate Lord, hired by rajas)\textdagger} --- A pirate lord with a fleet of painted dhows.
  \hfill \textit{He has nine wives; each is from a different kingdom. The ninth is a spy. He knows. He keeps her for the gossip — and because she is the only one who knows the delta's secret channels.} \texttt{[AUTHORITY]} \texttt{[PIRATE]}
\item[A] \textbf{The First Freedman (Legend, Sidhi)*} --- A slave who bought his freedom with a story. He now sits at the bottom of the Ninth Well.
  \hfill \textit{If you tell him a true story, he will tell you the location of a hidden lighthouse. The lighthouse is not on any map. It is where the rajas and the Sidhi first swore a secret pact.} \texttt{[LEGEND]}
\end{enumerate}

\section*{Clubs — Complications/Threats}
\label{sec:brasscoast-complications}

\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Pepper Dust Fire (Warehouse)\textdagger} --- A spark ignites a native-owned warehouse; the air is flammable. \texttt{[DISASTER]}
  \hfill \textit{The workers are trapped. Save them or save the cargo. The fire was set by an Admiralty agent to frame the Sidhi. The rajas are watching.}
\item \textbf{Admiralty Press Gang\textdaggerdbl} --- Kahfagian marines grab bystanders for the fleet. \texttt{[AUTHORITY]} \texttt{[COMBAT]}
  \hfill \textit{The marines are drunk. They do not recognise rank. One of the bystanders is a raja's son in disguise. The hills will demand blood if he is taken.}
\item \textbf{Green Lantern Lit (Sidhi Tavern)*} --- Someone lights the forbidden lantern. The dead come. \texttt{[DEATH]} \texttt{[FEAR]}
  \hfill \textit{They are looking for the person who killed them. That person is a native tax collector who sold out his village. The crowd will tear him apart — or protect him. Choose.}
\item \textbf{Shipbreaker's Curse (Admiralty Yard)\textdaggerdbl} --- A cursed ship in the yard breaks its chains. It wants a new victim — a ship's captain. \texttt{[DEATH]} \texttt{[COMBAT]}
  \hfill \textit{The captain is an Admiralty loyalist. The ship was a slaver that carried native prisoners. The ghosts want his blood. Do you help him or let them take him?}
\item \textbf{Toll Increase (Admiralty Decree)\textdaggerdbl} --- The port-master doubles duties on all goods. The increase is a bribe. \texttt{[AUTHORITY]} \texttt{[WEIGHT]}
  \hfill \textit{The bribe is to the Black Bishop. The Bishop has already accepted. Your protest is treason. The native merchants are watching to see which side you choose.}
\item \textbf{Coral Reef Piracy (Native-hired)\textdagger} --- Pirates hide behind the reef; they attack at dusk. \texttt{[COMBAT]} \texttt{[STEALTH]}
  \hfill \textit{The pirates are hired by a raja to disrupt Admiralty supply ships. They will not harm Sidhi vessels — but they cannot always tell the difference.}
\item \textbf{Elephant Rampage (Native Temple)\textdagger} --- A temple elephant breaks free and runs through the market. \texttt{[DISASTER]}
  \hfill \textit{It is looking for its mahout, who was arrested by the Admiralty for smuggling. The mahout is innocent. The elephant will not hurt children. Admiralty marines are another matter.}
\item \textbf{Smuggler's Betrayal (Neutral)} --- A smuggler captain sells the party to the Admiralty. \texttt{[SOCIAL]} \texttt{[TRAP]}
  \hfill \textit{The betrayal was a test. The captain wants to see if you can escape. If you do, you are worthy of trust — and the rajas will contact you.}
\item \textbf{Sea King's Tax (Pirate Blockade, hired by rajas)\textdagger} --- A Kalingan pirate fleet blockades the harbour. The tax is one ship. \texttt{[AUTHORITY]} \texttt{[COMBAT]}
  \hfill \textit{Choose which ship to give up. The Sea King will remember your choice. The crew will remember longer. The rajas are testing your loyalty.}
\item[J] \textbf{Drowned Priest's Demand (Neutral)} --- The ghost appears during the day. He asks for a name. \texttt{[DEATH]} \texttt{[SOCIAL]}
  \hfill \textit{The name must be someone who owes a death — a slaver, a corrupt customs officer, a raja who sold his own people. If you cannot give one, he takes yours.}
\item[Q] \textbf{Black Bishop's Verdict (False)\textdaggerdbl} --- The judge declares the party's cargo forfeit. The verdict is false. \texttt{[JUSTICE]} \texttt{[TRAP]}
  \hfill \textit{The bishop is bribed. The briber is the Admiralty. Prove it, or lose everything. The real verdict is written in disappearing ink. The rajas' spies are in the gallery.}
\item[K] \textbf{Sea King's Ninth Wife (Spy, works for the Veiled Queen)\textdagger} --- The spy reveals herself. She offers a deal: the Sea King's treasure for safe passage. \texttt{[POLITICS]} \texttt{[MAGIC]}
  \hfill \textit{The treasure is a curse. The curse is a fragment of the Kalrantha amulet. The amulet whispers in your sleep. It tells you which faction will betray you first.}
\item[A] \textbf{Undercliff Collapse (Smuggler's Cove)\textdagger} --- A rockslide seals the smugglers' cove. Inside are the party and a monster. \texttt{[DEATH]} \texttt{[TRAP]}
  \hfill \textit{The monster is a salt-crocodile. It is blind. It hunts by sound. The walls are listening — and so are the rajas' spies, who will arrive at dawn to see who survived.}
\end{enumerate}

\section*{Diamonds — Rewards/Leverage}
\label{sec:brasscoast-rewards}

\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Lighthouse Charter (Sidhi)*} --- The right to anchor in any Sidhi port without inspection. \texttt{[TRAVEL]}
  \hfill \textit{The key fits a lock on a sea-chest, not a harbour gate. The chest contains a map of smuggler's coves — and the names of rajas who pay protection.}
\item \textbf{Lighthouse Keeper's Lens (Sidhi)*} --- Look through it to see the nearest slaver ship. \texttt{[JUSTICE]}
  \hfill \textit{The lens also shows the faces of the slaves on board. They look back. They do not blink. It also shows the brand on the slaver's captain — and which raja sold them.}
\item \textbf{Freedman's Seal (Sidhi)*} --- Proof of emancipation. Glows when a slaver ship is near. \texttt{[SOCIAL]}
  \hfill \textit{It also glows when a Black Hand agent is near — and when a native informer is in the room. Trust the glow. It has never been wrong.}
\item \textbf{Pepper Merchant's Seal (Native)\textdagger} --- A wax seal from a dead factor. Present it to sell pepper at double price. \texttt{[TRADE]}
  \hfill \textit{The factor's ghost haunts the warehouse. He will ask for a story. The story must be about a bad deal — preferably one that cheated the Admiralty.}
\item \textbf{Salt Widow's Blessing (Native)\textdagger} --- A bag of bitter salt. Throw it into the sea to calm a storm. \texttt{[WEATHER]}
  \hfill \textit{The salt is from the widows' pans. They have blessed it with their tears. The storm will pause for one hour — long enough for a raja's fleet to escape.}
\item \textbf{Admiral's Pardon (Kahfagian)\textdaggerdbl} --- A signed document. Cancels one crime on the Brass Coast. \texttt{[AUTHORITY]}
  \hfill \textit{The Admiral signs with a glass eye. The signature is a lie. The pardon still works — but the eye watches you afterwards. It reports to the Admiralty's spymaster.}
\item \textbf{Sidhi Court Ruling (Sidhi)*} --- A written judgment. Win any maritime dispute in a Sidhi court. \texttt{[JUSTICE]}
  \hfill \textit{The ruling is valid only if witnessed by a lighthouse keeper. The keeper is Idris. He will witness for a story — and the story will reach the rajas within a day.}
\item \textbf{Elephant's Tooth (Native)\textdagger} --- A tusk from a temple elephant. Present it to any temple for shelter. \texttt{[SOCIAL]}
  \hfill \textit{The elephant was executed for killing an Admiralty tax collector. The temple will not refuse. They will also not ask questions. They will offer you a cup of palm wine — and a message from the Veiled Queen.}
\item \textbf{Shipbreaker's Hook (Admiralty)\textdaggerdbl} --- The hook of a dead shipbreaker. Use it to command the cursed ship in the yard. \texttt{[MAGIC]}
  \hfill \textit{The ship will obey one command — then it will try to drown you. The hook is warm. It remembers its owner's last words: ``Burn the toll ledgers.''}
\item[J] \textbf{Smuggler's Map (Neutral)} --- A chart of hidden coves, tattooed on a dead sailor's back. \texttt{[TRAVEL]}
  \hfill \textit{The skin is preserved in salt water. It must be kept wet or the map fades. The sailor's ghost will guide you — for a fee. The fee is a promise to avenge him against the Admiralty.}
\item[Q] \textbf{Sea King's Pardon (Native-hired)\textdagger} --- A flag of truce. Fly it to pass Kalingan waters unharmed. \texttt{[TRAVEL]}
  \hfill \textit{The flag is a shirt. The shirt belonged to a dead pirate — a native hero who fought the Admiralty. He may want it back. He will ask politely. If you refuse, the rajas will remember.}
\item[K] \textbf{Green Lantern's Flame (Sidhi)*} --- A small lantern that burns green. Light it to summon the dead. \texttt{[DEATH]}
  \hfill \textit{They will answer one question. Then they will ask one. The question is never easy. The answer is never free — but the dead know which faction is lying.}
\item[A] \textbf{Black Bishop's Ring (Covenant)} --- Wear it to be immune to Admiralty arrest in Sidhi ports. \texttt{[JUSTICE]}
  \hfill \textit{The nail is from a freed slaver. The slaves' ghosts protect you. They also watch. They judge. They remember. If you betray the rajas, the ring will grow cold — and the ghosts will abandon you.}
\end{enumerate}

\subsection*{Quick use notes}
\label{sec:brasscoast-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 \(\rightarrow\) 4, 6--10 \(\rightarrow\) 6, J/Q/K \(\rightarrow\) 8, A \(\rightarrow\) 10).
\item Diamonds are codified outcomes (charters, lenses, seals) that change position rather than call for a roll.
\item If any A appears, echo \textbf{coast-omens}---green flames that burn blue, salt that tastes sweet, a drowned face in a still puddle, a tiger's roar from the hills.
\item When a Club is drawn, tick one of the three faction timers (Admiralty Grip, Sidhi Ropeway Control, or Raja's Rebellion) based on the nature of the complication.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:brasscoast-mechanics}

\begin{itemize}
\item \textbf{The Three Faction Clocks (see above).} These are the campaign's pressure gauge. Use them to determine which faction acts first when the party dithers.
\item \textbf{The Sidhi Diaspora Network:} Sidhi characters gain +1 die on social rolls with other Sidhi, but --1 die on social rolls with Admiralty officials. With native rajas, they are neutral (no modifier) — but the rajas will always treat them as ``guests,'' never as allies.
\item \textbf{Native Hill Networks:} Native characters gain +1 die on social rolls with rajas and hill folk, but --1 die on social rolls with Admiralty officials. With Sidhi, they are cautious (no modifier) — the Sidhi are useful, but they are still foreigners.
\item \textbf{The Toll Clock [6] (renamed Admiralty Grip).} Already covered above.
\item \textbf{Green Flame Justice:} Sidhi lighthouse courts operate on the principle of ``witnessed truth.'' Any party member may request a green flame judgment instead of a conventional trial. Roll Presence + Lore (DV 4). On success, the judge rules in your favour; on a miss, the green flame shows a hidden crime you committed — and the penalty is doubled. However, if the crime was committed against a native, the rajas will learn of it.
\item \textbf{Ropeway Leverage:} The ropeways are the arteries of the coast. A character who controls a ropeway (by owning a share, bribing a foreman, or holding a Sidhi charter) may impose a ``ropeway toll'' on any cargo moving through that district: +1 Supply to the party or --1 Supply to a rival. Using this leverage ticks the Sidhi Ropeway Control timer by 1.
\item \textbf{Native Smuggling Routes:} A character with a native contact (e.g., a pepper factor or a salt widow) can use native smuggling routes to move goods without Admiralty inspection. This ticks the Raja's Rebellion timer by 1 but also ticks the Admiralty Grip timer by 1 if discovered.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Green Keeper} — The Witness. \textit{The flame that shows the truth, even when it burns — especially when it burns a foreigner's lie.}
    \item \textbf{The Hook} — Maelstraeus. \textit{The price of a second chance; the weight on a freedman's soul. The rajas say the Hook is Admiralty silver.}
    \item \textbf{The Tide Father} — Khemesh. \textit{The pressure beneath the keel; the toll that the sea collects. The widows say he is a native god, stolen by the Sidhi.}
    \item \textbf{The Lantern Bearer} — The Traveler. \textit{The light that guides lost ships — and lost rebellions. The Veiled Queen invokes him before every raid.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{Sidhi Freedmen and the Breath-less (Ashaan):} The Sidhi of the Brass Coast have family who were taken as slaves to Ashan. A freedman's seal can be used to smuggle a Breath-less agent out of Galanina. Captain Idris has done this nine times. The tenth may be his last — and the rajas are watching to see if the Sidhi are still loyal to their own.
  \item \textbf{Spice Trade and the Fhara (Way of Silk):} Pepper from the Brass Coast is a major commodity on the Fhara caravan routes. A Fhara well-judge may accept a Sidhi lighthouse charter as proof of credit. The charter must be witnessed by a living flame. The rajas have begun forging their own charters — and the Fhara cannot tell the difference.
  \item \textbf{The Undercliff and the Kalrantha Amulet (Dhahara):} The smuggler's cove is a known transit point for artifacts stolen from the Himadri temple. The Sea King's ninth wife claims to have seen the amulet pass through. She also claims the amulet spoke to her. It asked for a drink of water. The Veiled Queen believes the amulet is the key to driving out the Admiralty — but only if she can control it.
  \item \textbf{The Drowned Priest and the Hollow (Eastern Realms):} The ghost who walks the shore at low tide is said to be a Oath-Keeper who lost his way. Aelaerem leave offerings of salt to keep him from following them home. The salt must be from the pans of the Silent Widows. Any other salt will attract him — and the rajas have started using him as a weapon, luring him toward Admiralty patrols.
  \item \textbf{The Green Flame and the Dreamer (Ikasha):} The Sidhi lighthouse lamps burn green — the color of a promise kept, a debt paid, a memory not yet drowned. Some Breath-less agents whisper that the first green flame was lit from a coal stolen from the Dreamer's brazier, before Ikasha was sealed. The Sidhi do not confirm this. They do not deny it. They only tend the flame and ask no questions.
  \item \textbf{The Serpent's Judgment (Isoka) at the Freedman's Seal:} When a Sidhi freedman receives their brass seal, they are made to cut a thread in the presence of a lighthouse keeper. The thread represents the slaver's claim. If the cut is clean, the freedman is declared unbound. If the thread frays, the keeper demands a name — the name of the slaver who still holds a memory of them. This ritual is older than the Covenant. Some say it was taught by the Serpent herself, before she turned away from Ashaan.
  \item \textbf{The Raven's Shadow on the Ropeways:} At low tide, when the brass lanterns gutter, a shadow sometimes passes over the ropeways — a shadow with the shape of a bird, though no bird flies at night. The Sidhi workers touch their freedman's seals and whisper a single word: ``Remember.'' They do not say whose shadow it is. The Brass Coast rajas say it is Ikasha's dream, still looking for a shore. The Admiralty calls it a trick of the light. The Sidhi know better.
\end{itemize}

\subsection*{Brass Coast Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Tide:} Toss a crust into the harbour. Ignore the first \emph{Pepper Dust Fire} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Anchor Chain:} Tie a sailor's knot in a mooring rope. Cancel \emph{Admiralty Press Gang} or take --1 die on your next escape roll (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Lighthouse:} Whisper a dead captain's name into the green flame. Gain +1 Effect when smuggling this scene, but mark \emph{Obligation} to the Sidhi — and the rajas will learn the name. \texttt{[SUPERSTITION]}
  \item \textbf{Tiger's Whistle:} Whistle three times toward the hills before a deal. If a tiger answers, the deal is honest. If silent, the other party is lying. (GM spends 1 SB to have a tiger answer falsely once per session.) \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:brasscoast-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Admiralty customs, Sidhi ropemen, or a raja's spy}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{warehouse guards, lighthouse keepers, or hill lookouts}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{smugglers, the Black Bishop's agents, or a tiger}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favourable tide}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Admiralty marines, Sidhi enforcers, or a raja's war party} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{green flash, rising tide, or a dust storm from the hills}.
\item \textbf{Foe Appears:} An \textbf{Admiralty press gang, a Sidhi claim-keeper, or a native elephant} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Admiralty cutters, Sidhi ropemen, or hill warriors} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes—\textbf{toll increase declared / green lantern lit / raja's rebellion begins early}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment (Admiralty Grip, Sidhi Ropeway Control, or Raja's Rebellion).
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Admiral Voss, the Veiled Queen, or the Black Bishop} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically—\textbf{harbour catches fire / lighthouse collapses / hill fort's beacon is lit}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice—\textbf{omen appears / blessing lost / curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives — often by revealing that a faction has already made a secret pact.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Brass Coast (Toll Law):} Scales tip without weight, schedules rewrite themselves, a port-master's hook stamps a nullity — and a raja's dagger appears on the counter.
\item \textbf{Brass Coast (Lighthouse Law):} Green flame flickers, salt crystals form on skin, a drowned face appears in a puddle — and a hill lookout's horn sounds twice.
\item \textbf{Brass Coast (Smuggler's Law):} Ropeways hum with whispers, a parrot speaks a secret, a hidden cove appears on no map — and a tiger's shadow crosses the moon.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Sail):} On a hit, shift one band. On a strong hit, also impose \emph{Green Flash} (the lighthouse blinds pursuers).
\item \textbf{Cut the Toll (Presence + Sway):} On a hit, freeze range for one exchange; if a lighthouse charter is in your possession, +1 die.
\item \textbf{Weather the Raid (Resolve + Spirit):} On a hit in \emph{Admiralty Press Gang}, you pick who is taken; on a miss, the GM does — and the rajas will remember who you abandoned.
\end{itemize}

\subsection*{Boss Seeds (Toll \& Flame \& Rebellion)}
\begin{itemize}
\item \textbf{The Hook-Master (Admiralty Spymaster)\textdaggerdbl} — \emph{Phases}: Toll Forgeries \(\rightarrow\) Orchestrated Strike \(\rightarrow\) Port Coup. \emph{Cracks}: recover a true stamp, expose a forged schedule, survive a night in a lighthouse — or turn a raja against him.
\item \textbf{The Drowned Admiral (Ghost Fleet, Sidhi legend)*} — \emph{Phases}: Lost Ships Called Home \(\rightarrow\) Procession of Empty Lanterns \(\rightarrow\) Tide-Baptism of a Factor. \emph{Cracks}: relic green flame, survivor testimony, the lighthouse keeper's lens turning clear — and a native prayer spoken over the drowned.
\item \textbf{The Veiled Queen's Ascension (Native Rebellion)\textdagger} — \emph{Phases}: Whisper in the Mangroves \(\rightarrow\) The Tiger's Claw \(\rightarrow\) The Burning of the Schedules. \emph{Cracks}: unmask her face (revealing a childhood scar), expose her alliance with the Ashaani, or convince her that freedom requires mercy, not just blood.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
On the Brass Coast, power flows through tolls, ropeways, and the loyalty of the hill folk. The Admiralty holds the ports, the Sidhi hold the ropes, and the rajas hold the land. No single faction can win — but any faction can light the fuse.

\textbf{For the GM:}  
Use the three faction timers as your campaign's heartbeat. Every session, tick at least one timer based on the party's actions (or inaction). Let the players see the powder keg growing. When a timer fills, something explodes — a rebellion, a strike, a massacre. The party's job is not to prevent the explosion — it's to survive it and choose who rises from the ashes.

\noindent
\textbf{Native Focus:}  
The rajas are not victims — they are players. They have their own spies, their own fleets (via the Sea King), and their own magical traditions (tiger spirits, salt widows, elephant oracles). Whenever the party interacts with the Brass Coast, ask: ``How would a native see this?'' The Sidhi are refugees; the Admiralty are occupiers. The rajas are the hosts — and they are tired of guests.
\end{tcolorbox}

% Index entries
\index{Brass Coast|see{Kahfagian Dhahara, Three Powers}}
\index{Theme generator}
\index{Brass Coast generator}
\index{Places!Brass Coast}
\index{People!Brass Coast}
\index{Complications!Brass Coast}
\index{Rewards!Brass Coast}
\index{Sidhi diaspora}
\index{Admiralty}
\index{Native Dhaharan Rajas}
\index{Raja Govind}
\index{Veiled Queen}
\index{Three Faction Clocks}
\index{Toll Clock}
\index{Ropeway Control}
\index{Raja's Rebellion}
\index{Local Patrons!Brass Coast}