\section{Lethai-ar — ``The Crimson Valley''}
\label{chap:lethai-ar}

\textbf{Key Demographic: } Lethai-ar
\textbf{Capital City: } Ananse

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The bridge remembers every step. The silk knows every promise.}

\begin{quote}
  \textbf{Silk-Warden (Elite):}  
  ``The web-law is not written on parchment. It is woven into every bridge, every strand, every knot. When a promise is made, the valley remembers. When a promise is broken, the silk tightens. We are not judges. We are the loom. The spider spins; we only follow the thread.''
\end{quote}

\begin{quote}
  \textbf{Mark-Cutter (Commoner):}  
  ``Each line on my skin is a promise I witnessed. Each line is a story I carry. When I die, they will cut these marks from my body and bury them at the Ukwe's roots. Then the tree will remember what I witnessed. I hope it is enough. The thread does not break; it only passes to the next hand.''
\end{quote}

\textbf{Genre / Mood:} High-fantasy legal drama, spidersilk noir, ancestral witness.

\textbf{Starting Location:} The central silk bridge, where the gorge is deepest and the wind carries the smell of rust. A Silk-Warden blocks your path, her mask painted with roots: ``The web-law demands a toll. Not in coin – in a promise. Swear that you will not use this valley to carry steel to Ashaan, or turn back now. The thread is watching. So is the spider.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Crimson Valley is a deep jungle rift between Sekogo and Oshiira – a wound of red stone, carved by a river that runs rust‑colored after every rain. Here, the Lethai-ar guard their borders with fierce neutrality, their chestnut‑brown skin blending with the canyon walls. Silk bridges span the gorges, woven from spider‑sap and oath‑hair, strong enough to hold a caravan and light enough to sway in the monsoon wind. The web-law is unwritten, carried in the memory of each warden; it keeps both empires from using the valley as a shortcut to war.

Unlike their northern kin who fell from grace, the Lethai-ar are adopted children of the old powers – not fallen, but chosen to carry memory. Each can name the specific mountain, river, or grove where their ancestor left the spirit world to become mortal. Their skin‑tattoos (Marks) are woven with red ochre and jungle sap, each line a promise witnessed, each curve a story recorded. When a Lethai-ar dies, their Marks are cut from the skin and buried at the Ukwe's roots, so the tree may remember what they witnessed.

The valley itself is alive. The silk remembers every footfall; the river whispers the names of those who have crossed; and in the oldest bridges, the spider – Inaea, the weaver of bonds – still spins. The Lethai-ar do not worship her, but they honor the thread. They say that every promise is a strand in her web, and that a broken oath sends a tremor through the whole valley.

\vspace{2mm}
\textbf{What you notice first:} The whisper of silk ropes in the wind. The smell of red ochre and damp stone. The way every Lethai-ar touches their left wrist when they speak a name – a gesture of witness.

\textbf{The one rule that kills newcomers:} Never step onto a silk bridge without first naming a witness. The bridge will remember your footsteps. If you fall, it will ask why no one was watching.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of Witness}

Power here flows from witnessed promises, silk strands, and the web-law's quiet authority. The valley is neutral ground – Oshiiran grain barges do not float here, and Ashaani armies dare not march. The Lethai-ar trade silk for honey with Sekogo druids and stonesong for poison with Ngomebe masons. Their Marks are not decoration; they are a public memory, written on skin, read by anyone who knows the phase of the crimson moon. A Lethai-ar who bears a Mark they cannot fulfil loses standing; a warden who cuts a bridge without cause is exiled. But the true authority is the web itself – the invisible threads that connect every oath ever sworn in the valley. The Lethai-ar do not create the web. They tend it.

\subsection*{Prominent NPCs of the Crimson Valley}
\label{sec:lethaiar-npcs}

\begin{tabular}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Warden Anara & Silk‑warden of the central span & Keep the Oshiiran grain barges from crossing & Her left arm is covered in Marks from wrist to shoulder. She will not say what they mean. \\
\hline
Mark‑Cutter Tebe & Tattooist who records promises & Witness a treaty between Sekogo and Ngomebe & She uses a needle of silk‑fiber and ink made from crushed crimson stone. The ink glows under the full moon. \\
\hline
Bridge‑Keeper Solan & Guardian of the eastern crossing & Prevent Ashaani scouts from using the valley & He has not slept in nine days. He says the bridge whispers when someone lies. \\
\hline
Elder Mira & Memory‑keeper of the valley's spirits & Record the name of every ancestor who walked the spirit path & She carries a bundle of silk cords, each tied to a different knot. The knots are the names. \\
\hline
The Spider's Child & Tulkani trader of oath‑silk (whisperer of Inaea) & Buy a bolt of crimson silk for a story – and to listen & She is not Lethai-ar but has lived in the valley for thirty years. The wardens tolerate her. She never closes her eyes. \\
\hline
The Unmarked & A Lethai-ar who refused any Marks & An outcast who lives in the gorge below the bridges & Her skin is blank. She is invisible to the web‑law – and to the valley's spirits. She carries a single knotted thread. \\
\hline
\end{tabular}

\subsection*{Names of the Crimson Valley}
\label{sec:lethaiar-names}

Lethai-ar names are soft, with rolling consonants and vowel endings. Surnames often denote a bridge, a gorge, or a spirit‑ancestor. Epithets refer to Marks or witnessed deeds.

\begin{tabular}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Solan & Anara & Silkweave & \textit{the Unbroken} \\
\hline
Tebe & Mira & Gorge‑born & \textit{the Ninth Witness} \\
\hline
Liam & Nira & Root‑memory & \textit{the Marked} \\
\hline
Veth & Liana & Spider‑kin & \textit{the Bridge‑Heart} \\
\hline
Renn & Sora & Crimson‑touch & \textit{the Unweaver} \\
\hline
Thorn & Aria & Oath‑strand & \textit{the Still Watcher} \\
\hline
\end{tabular}

\paragraph*{(Bridge/Gorge/Roots)} Silk bridge spanning a crimson gorge – ropes sing in the wind; a Lethai-ar village built into the canyon wall; the Ukwe's roots clutching the water below the central span.

\section*{Spades — Places}
\label{sec:lethaiar-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Central Silk Bridge} – A wide span woven from crimson and gray strands. The bridge hums when a promise is made on it.
  \hfill \textit{The pitch tells the warden if the promise is true. If the thread vibrates at the ninth frequency, the spider is listening.} \texttt{[WITNESS]}
\item \textbf{Mark‑Cutter's Studio} – A cave with a stone table and pots of ochre.
  \hfill \textit{The walls are carved with Marks that were never cut into skin. They are stories the cutter refused to witness. The stone is warm to the touch.} \texttt{[CRAFT]}
\item \textbf{Ukwe's Riverside Root} – A massive root that dips into the crimson river.
  \hfill \textit{The root is warm. Touch it, and the tree will show you one forgotten promise from a dead Lethai-ar. The water ripples in response.} \texttt{[MEMORY]}
\item \textbf{Abandoned Silk Span} – A broken bridge hanging over the gorge. The cut ends are frayed.
  \hfill \textit{The cut was not clean – someone broke an oath mid‑crossing. The strands still twitch.} \texttt{[DEATH]}
\item \textbf{Spirit‑Anchor Stone} – A boulder where the first ancestor left the spirit world.
  \hfill \textit{The stone is smooth. It is warm even in winter. Lethai-ar leave offerings of red ochre here – and sometimes a single knotted thread.} \texttt{[RITUAL]}
\item \textbf{Warden's Watchtower} – A stone needle built into the canyon wall, with a signal mirror.
  \hfill \textit{The mirror flashes in code. The code is the web‑law. No outsider has broken it. The code changes when the spider dreams.} \texttt{[AUTHORITY]}
\item \textbf{Hollow of the Unmarked} – A cave below the eastern bridge, hidden by a waterfall.
  \hfill \textit{The cave has no silk. The Unmarked lives here, weaving nets from her own hair. Her hair is knotted.} \texttt{[MYSTERY]}
\item \textbf{Crimson River Ford} – A shallow crossing where the water is clear, not red.
  \hfill \textit{The ford is sacred. No oath sworn here can be broken. The water would turn black – and the spider would feel it.} \texttt{[OATH]}
\item \textbf{Elder's Cord‑House} – A longhouse where the memory‑knots are stored.
  \hfill \textit{Each cord is a name. The cords are hung from the ceiling. They move in no wind. When a name is forgotten, a cord falls.} \texttt{[MEMORY]}
\item[\card{\spadesuit}{J}] \textbf{The Spider's Stall} – A Tulkani merchant's tent at the valley's southern exit. She sells silk to anyone.
  \hfill \textit{The price is a story. The story must be about a thread that held – or a thread that broke. She never blinks.} \texttt{[TRADE]}
\item[\card{\spadesuit}{Q}] \textbf{Bridge of the Ninth Strand} – A legend: a bridge woven from a single hair.
  \hfill \textit{It appears only at dawn on the first day of spring. Those who cross it gain a new name – and a new thread to their web.} \texttt{[LEGEND]}
\item[\card{\spadesuit}{K}] \textbf{The Cut Stone} – A boulder split by an oath‑breaker's fall.
  \hfill \textit{The split is shaped like a human body. The stone weeps water when the oath‑breaker's name is spoken. The water tastes of rust.} \texttt{[DEATH]}
\item[\card{\spadesuit}{A}] \textbf{The First Bridge (Myth)} – Where the first Lethai-ar wove the first strand.
  \hfill \textit{It is invisible. Only those who have witnessed nine true promises can see it. The spider sits at the far end, waiting.} \texttt{[MYTH]}
\end{enumerate}

\paragraph*{(Warden/Cutter/Trader)} Silk‑warden with a mask painted with the Ukwe's roots; Mark‑cutter with a needle of silk‑fiber; Tulkani spider‑trader with a bundle of oath‑cords.

\section*{Hearts — People \& Factions}
\label{sec:lethaiar-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Silk-Warden} – Mask carved from crimson wood, painted with oath‑names.
  \hfill \textit{She never blinks when standing on the bridge. The bridge does not forgive distraction. She counts her breaths in eights.} \texttt{[AUTHORITY]}
\item \textbf{Mark-Cutter} – Hands stained red with ochre, eyes that measure skin.
  \hfill \textit{She asks each client: ``What promise do you want the world to remember?'' The answer becomes a Mark – and a thread in the web.} \texttt{[CRAFT]}
\item \textbf{Bridge-Keeper} – Old, gray‑haired, missing three fingers from a snapped strand.
  \hfill \textit{He carries a whistle made from a spider's fang. The whistle calls the bridge to tighten – or to loosen. He never uses it.} \texttt{[WITNESS]}
\item \textbf{Elder (Memory‑Keeper)} – Bundles of silk cords hang from her belt.
  \hfill \textit{She knows every name carved into every root of the Ukwe. She has forgotten her own. The cords remember it for her.} \texttt{[MEMORY]}
\item \textbf{Tulkani Spider (Trader)} – Wears a shawl of crimson silk, speaks in riddles.
  \hfill \textit{She pays for silk with stories. The stories become new strands. The bridge grows. She never closes her eyes – she says the spider does not sleep.} \texttt{[TRADE]}
\item \textbf{The Unmarked} – A Lethai-ar without Marks, living in the hollow.
  \hfill \textit{She whispers to the waterfall. The waterfall whispers back. No one knows what they say. Her thread is her own.} \texttt{[MYSTERY]}
\item \textbf{Youth Seeking a Mark} – A young Lethai-ar on their first witness journey.
  \hfill \textit{They carry a single strand of undyed silk. When they witness a promise, they will dye it crimson. The dye is their own blood.} \texttt{[COMMUNITY]}
\item \textbf{Oath‑Breaker's Ghost} – A translucent figure pacing the broken bridge.
  \hfill \textit{It repeats the words of the promise it broke. The words are getting quieter. When they stop, the thread will snap.} \texttt{[DEATH]}
\item \textbf{Sekogo Honey Trader} – A druid with a jar of dark honey, trading for silk poison.
  \hfill \textit{He offers the honey as a gift. The wardens test it by feeding a drop to the bridge. The bridge sways if it is poisoned – or if the gift is false.} \texttt{[TRADE]}
\item[\card{\heartsuit}{J}] \textbf{Ngomebe Stone‑Singer} – A mason with a tuning fork, here to buy silk for sledge‑wax seals.
  \hfill \textit{He taps the bridge with his fork. The note tells him how many promises it has witnessed. The note is always an octave below the spider's hum.} \texttt{[CRAFT]}
\item[\card{\heartsuit}{Q}] \textbf{The First Ancestor's Echo} – A voice heard at the Spirit‑Anchor Stone at midnight.
  \hfill \textit{It asks: ``Why did you leave the spirit world?'' The answer must be a promise. The spider listens.} \texttt{[RITUAL]}
\item[\card{\heartsuit}{K}] \textbf{The Spider's Daughter} – A half‑Lethai-ar child raised by the Tkanki trader. She has no Marks.
  \hfill \textit{Her skin is blank. She is learning to read knots – and to spin her own thread. She never asks twice.} \texttt{[COMMUNITY]}
\item[\card{\heartsuit}{A}] \textbf{The Ninth Warden (Legend)} – A warden who cut her own bridge to stop an invasion.
  \hfill \textit{She is said to walk the valley at dusk, carrying a coil of silk. She offers one free crossing – but the crossing leads to the spider's web.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Strand/Challenge/Fall)} A silk strand snaps – lose 1 Supply to repair; a Lethai-ar challenges you to recite the web‑law; a traveler falls from the bridge – the party must witness their last words.

\section*{Clubs — Complications/Threats}
\label{sec:lethaiar-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Snapped Strand} – A rope on a bridge breaks. Lose 1 Supply to repair.
  \hfill \textit{The break is clean. Someone cut it with a blade. The blade was from Ashaan – or from a broken promise.} \texttt{[DISASTER]}
\item \textbf{Web-Law Challenge} – A warden blocks the bridge: ``Recite the ninth clause.''
  \hfill \textit{The ninth clause is unwritten. Each warden carries a different version. You must guess. The spider knows the true one.} \texttt{[SOCIAL]} \texttt{[RIDDLE]}
\item \textbf{Fall from the Bridge} – A traveler slips. The bridge asks: ``Who witnessed this?''
  \hfill \textit{If no one answers, the traveler's name is erased from the Ukwe's roots – and the spider weaves a new strand.} \texttt{[DEATH]} \texttt{[WITNESS]}
\item \textbf{Oshiiran Surveyors} – Grain auditors arrive with a map, claiming the valley's water rights.
  \hfill \textit{They carry a treaty stamped with the Widow's seal. The seal is a forgery. The bridge hums in warning.} \texttt{[AUTHORITY]} \texttt{[POLITICS]}
\item \textbf{Ashaani Scouts (Disguised)} – Veiled Sister agents posing as spice merchants.
  \hfill \textit{They do not touch their left wrist when speaking a name. The wardens notice. The thread tightens.} \texttt{[STEALTH]} \texttt{[POLITICS]}
\item \textbf{Mark-Cutter's Witness Called} – A Lethai-ar's Marks begin to glow red. A promise is overdue.
  \hfill \textit{The witness is not present. The party is the closest witness. The duty transfers – as does the thread.} \texttt{[WITNESS]} \texttt{[COMMUNITY]}
\item \textbf{Bridge Poisoning} – Someone smears a solvent on the silk strands. The bridge begins to fray.
  \hfill \textit{The solvent is made from Sekogo honey and Ashaani dust. The Spider knows who sold it. The threads are weeping.} \texttt{[MAGIC]} \texttt{[DISASTER]}
\item \textbf{Elder's Memory Fails} – The cord‑keeper forgets a name. The knot unravels.
  \hfill \textit{The forgotten name is an old promise. The Ukwe's root trembles. Something is coming – or someone.} \texttt{[MEMORY]} \texttt{[DEATH]}
\item \textbf{The Unmarked's Bargain} – The outcast offers to guide the party through the gorge. Her price is a promise.
  \hfill \textit{The promise: ``Do not speak my name after I die.'' Refuse, and she leaves you to the spiders. Accept, and she ties a knot.} \texttt{[SOCIAL]} \texttt{[WITNESS]}
\item[\card{\clubsuit}{J}] \textbf{Crimson River Floods} – Rain turns the ford to a torrent. The bridge is the only crossing.
  \hfill \textit{The water is red. Faces appear in the foam. They are oath‑breakers, still trying to cross. The spider watches.} \texttt{[WEATHER]} \texttt{[DEATH]}
\item[\card{\clubsuit}{Q}] \textbf{Warden Schism} – Two wardens disagree on the web‑law. They cut each other's bridges.
  \hfill \textit{The valley is now divided. The party must choose which side to cross – or stay stuck forever. The spider waits.} \texttt{[AUTHORITY]} \texttt{[POLITICS]}
\item[\card{\clubsuit}{K}] \textbf{The First Bridge Appears} – The mythical bridge manifests at dawn. Crossing it gives a new name.
  \hfill \textit{The new name is a promise. The promise is to the First Ancestor. The ancestor will collect – and the spider will weave.} \texttt{[MYTH]} \texttt{[MAGIC]}
\item[\card{\clubsuit}{A}] \textbf{The Ninth Warden Returns} – The legend appears, offering a free crossing.
  \hfill \textit{The crossing leads to the spider's web. The Hollow asks: ``What promise did you break?'' Answer poorly, and you become a new strand.} \texttt{[DEATH]} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Cord/Mask/Shard)} Silk truth‑cord (enforce any oath); warden's mask (safe passage on any bridge for one crossing); Ukwe root shard (one question about a forgotten promise).

\section*{Diamonds — Rewards/Leverage}
\label{sec:lethaiar-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Silk Truth‑Cord} – A braided strand dyed crimson. Use it to enforce any oath.
  \hfill \textit{The cord tightens when the oath is broken. It will not stop until the promise is kept – or the spider releases it.} \texttt{[OATH]}
\item \textbf{Warden's Mask} – A wooden mask painted with roots. Safe passage on any bridge for one crossing.
  \hfill \textit{The mask has no eye holes. You must trust the bridge to see for you. The bridge sees through the spider's eyes.} \texttt{[TRAVEL]}
\item \textbf{Ukwe Root Shard} – A piece of the tree's root that grows near the river.
  \hfill \textit{Touch it, and the Ukwe will answer one question about a forgotten promise. The answer is always true – and always a thread.} \texttt{[MEMORY]}
\item \textbf{Mark‑Cutter's Needle} – A silk‑fiber needle. Use it to record a witnessed promise as a temporary Mark.
  \hfill \textit{The Mark fades after one season. The story remains – and the spider remembers.} \texttt{[CRAFT]}
\item \textbf{Bridge‑Keeper's Whistle} – A fang whistle. Blow it to tighten any silk bridge for one scene.
  \hfill \textit{The whistle works once. The bridge will remember the favor – and the spider will know.} \texttt{[AUTHORITY]}
\item \textbf{Elder's Memory Cord} – A knotted silk cord that contains a forgotten name.
  \hfill \textit{Untie the knot to learn the name. The name is a promise. Fulfil it, or the cord will retie itself – and the spider will watch.} \texttt{[MEMORY]}
\item \textbf{Spider's Story‑Strand} – A bolt of crimson silk bought with a story.
  \hfill \textit{The silk is soft. It glows under moonlight. Trade it for safe passage through Tkanki camps – or give it to the spider.} \texttt{[TRADE]}
\item \textbf{The Unmarked's Net} – A net woven from human hair.
  \hfill \textit{Throw it to catch a spirit. The net can hold a bound spirit for one hour. The spirit will bargain for freedom – with a promise.} \texttt{[MAGIC]}
\item \textbf{Crimson Ochre Pot} – A clay pot of red pigment. Use it to Mark a promise on stone or skin.
  \hfill \textit{The ochre never dries. The promise remains witnessable forever – until the spider cuts the thread.} \texttt{[CRAFT]}
\item[\card{\diamondsuit}{J}] \textbf{Ninth Strand Fragment} – A piece of the legendary bridge. Once per campaign, cross any gap as if on solid ground.
  \hfill \textit{The fragment crumbles after use. The first bridge will remember you – and the spider will ask for a toll.} \texttt{[LEGEND]}
\item[\card{\diamondsuit}{Q}] \textbf{The Cut Stone's Water} – A vial of water that seeps from the oath‑breaker's boulder.
  \hfill \textit{Drink it to remember a promise you broke. The memory is painful. The pain is the price – and the thread.} \texttt{[DEATH]}
\item[\card{\diamondsuit}{K}] \textbf{First Ancestor's Blessing} – A whispered promise from the Spirit‑Anchor Stone.
  \hfill \textit{Once, you may reroll any failed roll involving a witnessed promise. The ancestor watches – and the spider weaves.} \texttt{[RITUAL]}
\item[\card{\diamondsuit}{A}] \textbf{The Ninth Warden's Coil} – A coil of silk said to belong to the legendary warden.
  \hfill \textit{Use it to cut a bridge without angering the web‑law. The coil then unravels – and the spider's shadow falls on you.} \texttt{[LEGEND]}
\end{enumerate}

\section*{Quick use notes}
\label{sec:lethaiar-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2–5 \(\rightarrow\) 4, 6–10 \(\rightarrow\) 6, J/Q/K \(\rightarrow\) 8, A \(\rightarrow\) 10).
\item Diamonds are codified outcomes (cords, masks, shards) that change position rather than call for a roll.
\item If any A appears, echo \textbf{valley‑omens} – silk cords humming in still air, the crimson river running clear, a bridge that sways without wind. The spider is near.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:lethaiar-mechanics}

\begin{itemize}
\item \textbf{The Web‑Law:} The unwritten legal code of the Crimson Valley. Each warden carries a different clause in memory; together, they form a complete web. To cross a bridge, a traveler must answer a question about the web‑law. The answer changes with each warden. On a successful roll (Wits + Lore, DV 3), you may bypass the challenge; on a failure, you must offer a promise instead – and the spider will witness it.
\item \textbf{Marks (Skin‑Tattoos):} Each Lethai-ar's Marks are a public record of promises witnessed. The Marks are cut from the skin after death and buried at the Ukwe's roots. To read a living Lethai-ar's Marks, one must know the phase of the crimson moon and the position of the valley's spirits. A character who studies a Lethai-ar's Marks for a scene may ask the GM one question about a promise that Lethai-ar has witnessed; the GM must answer truthfully – but the spider may also listen.
\item \textbf{Spirit Ancestry:} Every Lethai-ar can name the specific spirit (mountain, river, grove) where their ancestor left the spirit world to become mortal. That spirit is their witness; they owe it a promise at birth, and that promise is never fully paid. Once per journey, a Lethai-ar may invoke their spirit ancestor to gain +1 die on a roll related to that spirit's domain (e.g., a river ancestor aids a crossing). The spider notes the favor.
\item \textbf{Witness Economy:} Instead of tracking debt, the Crimson Valley uses \textbf{Witness} as currency. A character who witnesses a promise gains 1 Witness token. Tokens can be spent to: gain +1 die on a roll involving that promise, force an oath‑breaker to reroll a successful deception, or demand a favor from someone who witnessed the same promise. Witness tokens are lost when the promise is fulfilled or broken – and the spider watches the exchange.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The First Weaver (Inaea):} The spider who taught the first knot; mercy is a strand, not a cut. Her web is the valley's memory.
    \item \textbf{The Serpent Judge (Isoka):} The fang that decides; a promise kept is a skin shed. She watches from the river stones.
    \item \textbf{The Crimson Witness (The Witness):} The flame that sees through silk; no oath hidden from its light. The bridge glows when it passes.
    \item \textbf{The Bridge Father (The Traveler):} The road that crosses the gorge; every crossing is a promise. He walks the strands at dusk.
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross‑Expansion Hooks)}
\label{sec:lethaiar-lore}

\begin{itemize}
  \item \textbf{Crimson Silk and the Way of Silk:} The silk of the Crimson Valley is prized across the continent for vow‑cords, well‑contracts, and bridge‑building. The Tkanki spider‑traders are the primary merchants; a single bolt can buy safe passage from Fhara to Dhahara. A Tkanki story‑seller may accept crimson silk as payment for a truth – but the spider will ask for a copy.
  \item \textbf{The Ukwe's Roots and Sekogo:} The Ukwe's roots extend into the Crimson Valley. The Lethai-ar bury their dead Marks at the riverbank root, creating a unique relationship with the Speaking Tree. The Ukwe knows every promise ever made in the valley. A Sekogo druid may ask the Ukwe to witness a new oath, binding it with the weight of the valley – and the spider's thread.
  \item \textbf{Web‑Law and the Covenant (Mykkiel):} The Covenant judges respect the web‑law as a valid legal system. A Lethai-ar truth‑cord is admissible in a Covenant court. Some judges have learned to read Marks. A Covenant justiciar may offer to translate a web‑law dispute into imperial law – but the translation always favors the spider.
  \item \textbf{The Unmarked and the Hollow (Eastern Realms):} Lethai-ar who refuse all Marks become invisible to the web‑law – and to the Hollow. The Hollow cannot track them. Some Aelaerem seek out the Unmarked to learn the secret of blank skin. The Unmarked charge a promise: ``Do not seek me again.'' The spider does not visit them.
\end{itemize}

\subsection*{Lethai-ar Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Bridge:} Crumble a crust onto the first strand. Ignore the first \emph{Snapped Strand} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Cord:} Tie a single knot in a truth‑cord. Cancel \emph{Web-Law Challenge} or take –1 die on your next negotiation (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Root:} Whisper a dead warden's name into the Ukwe's root. Gain +1 Effect when crossing a bridge this scene, but mark Obligation to the First Ancestor – and the spider will know. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:lethaiar-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from Silk‑Wardens or Mark‑Cutters.
\item \textbf{Noise:} Sounds of your actions alert nearby bridge‑keepers or Tkanki traders.
\item \textbf{Trace:} Evidence of your passage marks your route for Elders or the Unmarked.
\item \textbf{Delay:} A brief but meaningful setback costs you time or safe crossing.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant resource timer.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} Local Silk‑Warden or Bridge‑Keeper becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to snapped strand or rising river.
\item \textbf{Foe Appears:} An Oshiiran surveyor or Ashaani scout arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes Compromised/Neglected.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional Wardens, Mark‑Cutters, or bridge guards arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes – bridge cut / web‑law broken / First Bridge appears.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Elder, Ninth Warden, or First Ancestor's Echo} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically – \textbf{bridge collapses / river turns black / Ukwe root rises}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice – \textbf{omen appears / blessing lost / curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Crimson Valley (Web‑Law):} Silk strands glow, truth‑cords tighten, the Ukwe's root hums a name.
\item \textbf{Crimson Valley (Marks):} Tattoos itch, ochre stains spread, a dead Mark appears on living skin.
\item \textbf{Crimson Valley (Bridges):} Ropes whisper, the gorge echoes with footsteps that are not there, a bridge sways toward the Hollow.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Swaying Bridge}.
\item \textbf{Cut the Strand (Presence + Sway):} On a hit, freeze range for one exchange; if a truth‑cord is in your possession, +1 die.
\item \textbf{Weather the Crossing (Resolve + Spirit):} On a hit in \emph{Fall from the Bridge}, you pick who catches the fallen; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Silk \& Promise)}
\begin{itemize}
\item \textbf{The Unweaver (Nine Broken Strands)} – \emph{Phases}: Snapped Cords \(\rightarrow\) Orchestrated Falls \(\rightarrow\) Bridge Coup. \emph{Cracks}: repair a broken strand, expose a false witness, survive a crossing without a promise.
\item \textbf{The Silent Warden (Oath of the Unmarked)} – \emph{Phases}: Forgotten Promises \(\rightarrow\) Procession of the Blank \(\rightarrow\) Skin‑Baptism of an Elder. \emph{Cracks}: relic truth‑cord, survivor testimony, the Marks turning against them.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Lethai-ar, power flows through witnessed promises, silk strands, and the authority of the wardens. The web‑law is unwritten, but every warden carries a piece of it. Marks are public, stories are visible, and the Ukwe remembers everything. The spider – Inaea – does not rule, but she watches. Her threads are the bonds between families, the knots that hold the valley together, and the silent witness to every oath.

\textbf{For the GM:} Power in the Crimson Valley is measured in witnessed oaths and the trust of the bridge. Patronage often takes the form of truth‑cords, masks, or root shards – each tied to a specific promise. To emphasize this:
\begin{itemize}
\item Tie rewards to objects that record or enforce promises (cords, masks, shards) that can be challenged, broken, or called in.
\item Let rival wardens issue conflicting interpretations of the web‑law, forcing players to choose whose bridge to cross.
\item Use the bridges, Mark‑cutter's studio, and the Ukwe's root as arenas for social contests, where a witnessed promise is a weapon and a broken strand is a death sentence.
\item Remember that the spider is not a god – she is a presence. She does not speak, but the valley hums with her attention. When a promise is broken, the threads tremble.
\end{itemize}
In the Crimson Valley, the bridge remembers every step – and the spider forgets no promise.
\end{tcolorbox}

% Index entries
\index{Lethai-ar|see{Crimson Valley}}
\index{Theme generator}
\index{Crimson Valley generator}
\index{Places!Lethai-ar}
\index{People!Lethai-ar}
\index{Complications!Lethai-ar}
\index{Rewards!Lethai-ar}
\index{Silk-Warden}
\index{Mark-Cutter}
\index{Web-law}
\index{Truth-cord}
\index{Inaea (spider)}
\index{Ukwe (root)}
\index{Local Patrons!Lethai-ar}
