\section{Ecktoria --- ``Marble & Fire''}
\label{chap:ecktoria}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The flame remembers. The marble judges. The ledger never sleeps.}

\begin{quote}
  \textbf{Grand Magistrate (Elite):}  
  ``The law here is written not in books but in marble and flame. Every edict carries the weight of centuries, and every judgment echoes in the very stones of the Forum. But the Forum has grown quiet of late. The patricians no longer listen.''
\end{quote}

\begin{quote}
  \textbf{Ferrymaster (Commoner):}  
  ``I've seen the city burn three times and rise again like a phoenix made of stone. The marble remembers, the bells tell the truth, and the fire judges all — especially those who think coin can buy clean hands. But this time the fire isn't burning. It's \emph{smiling}. The rich have stopped looking at the flames. They only look at each other.''
\end{quote}

\textbf{Genre / Mood:} Imperial decay, mercantile noir, flame‑lit judgment, Renaissance decadence — a city walled off from its own rot, where the wealthy dance while the poor cough ash.

\textbf{Starting Location:} The Coin‑House floor, marble cracked with soot lines, where votive candles burn beside melted icons. A factor with smoke‑singed ledgers offers a purse. ``The Lame King moves,'' she says. ``And the Basilica burns with him. Your credit is good — for now. Find out who lit the fire, or the Flame will audit your soul at dawn. And stay away from the Amber Palace. The Vigil has been… \emph{smiling}.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Once the heart of the Utaran Empire, Ecktoria has worn its laurels into chains. Three centuries of sack, fire, and slow decline have not stripped its marble forums nor its gold‑sealed coin‑houses — but each day, more of its wealth slips into the hands of mercenaries and factions. The Black Banners march on its purse, avenging old Ykrul wounds by proxy. Processions still crown its stairways, yet ash lines every step. The city burns not only with memory, but with ambition still unquenched.

Yet the true sickness is not in the streets. It is in the \textbf{Amber Palace}, where the old patrician families have walled themselves away. They sit on laurels preserved from the empire, drinking wine that costs a village's harvest, whispering deals with merchants who have never seen a loom. \textbf{Maelstraeus}, the demon of commerce and greed, has sunk his hooks into their ledgers; every contract they sign compounds interest in blood. \textbf{Livaea}, the demon of seduction and desire, has wrapped their hearts in silk; they marry for influence, divorce for scandal, and love only the reflection of themselves in gilded mirrors. The common folk call them the \emph{Unburdened} — not because they carry no weight, but because they have forgotten what weight feels like.

The \textbf{Vigil} — once the empire's secret police, now a fractured ghost of an institution — fights a silent war over the dream of restoration. One faction, the \textbf{Chain-Lanterns}, serves the Everflame and the Synod, hunting heresy and latent magic. Another, the \textbf{Unburnt}, answers only to the old patrician houses, ensuring that no whisper of rebellion reaches the Amber Palace. A third, the \textbf{Red Vigil}, has sold itself to Maelstraeus, auditing debts and collecting interest in screams. They do not fight each other openly. They meet in the hypogeum of the Arena, under the gaze of statues that weep marble tears.

Beyond the capital's walls, the province is a patchwork of struggling towns, abandoned fortresses, and grainlands whose harvests are disputed by rival tax‑collectors. The Everflame Orthodoxy preaches unity, but each sermon is a coded signal for a political camp. The statues watch. The fire judges. And in the sewers, the Vigil's informers listen for the name of the next person to vanish.

\vspace{2mm}
\textbf{What you notice first:} The smell of old incense and fresh bread — and beneath it, the faint sweetness of rot. The way crowds part for a procession — and close again for a knife. The marble steps, worn smooth by centuries of petitioners. The way the wealthy never step on the cracks.

\textbf{The one rule that kills newcomers:} Never refuse a toast in a coin‑house. The ledgers are the true judges, and a declined cup is a declined credit. You will not borrow again. But more than that: never accept a gift from a patrician without naming its price aloud. Maelstraeus and Livaea both listen, and a gift accepted in silence is a contract signed in blood.
\end{tcolorbox}

\subsection*{Regional Mood: The Flame That Remembers — And The Dancers Who Forgot}

Here, the past does not fade — it is polished into monuments. Every statue, every obelisk, every triumphal arch is a legal precedent carved in stone. Ecktoria is not one city but three: the marble forums of the old patricians (where the Unburdened hold masques that last for weeks, their faces painted with ash from the Everflame), the soot‑stained alleys of the mercenary camps (where Black Banner captains drink with Vigil informers), and the basilicas where the Everflame's priests balance piety against profit (and sometimes sell indulgences for the sin of being born poor). The Grand Magistrate presides over a skeleton empire, but the real power moves in coin‑houses, grain ledgers, and the whispered judgments of the Censor's Hall — and in the dreams of patricians who no longer remember the scent of rain.

The \textbf{Vigil} does not patrol the streets. It \emph{permeates} them. Every tavern has an informer; every bathhouse has a spy; every marriage is witnessed by someone who reports to a faction of the Vigil. The Chain-Lanterns hunt magic; the Unburnt hunt sedition; the Red Vigil hunts \emph{debt}. They are not allies. They are not enemies. They are three creditors fighting over the same soul.

The \textbf{patricians} — the Old Blood — no longer govern. They \emph{perform} governance. A session of the Parlement might see two senators debate the price of grain while their slaves hold silver trays of honeyed figs. A decree might be passed that doubles the salt tax, then rescinded an hour later because a duchess sneezed. They are not cruel. They are \emph{distracted}. Livaea has taught them to crave the next pleasure before they have finished the current one. Maelstraeus has taught them that every transaction is a game, and every game is winnable if you have enough coin. They do not see the bread riots. They see a shortage of \emph{entertainment}.

The \textbf{Everflame} still burns in the basilicas, but its light has grown dim. The High Priest speaks of unity, but his sermons are interrupted by coughing from the pews — not from illness, but from the ash that falls like snow. The common folk whisper that the flame no longer judges truth; it only \emph{records} it, filing confessions in a vault that no one ever opens. The Vigil has a key. The patricians have a copy. The poor have nothing but their names, which the Censor's Hall writes in disappearing ink.

\subsection*{At a Glance}
\begin{itemize}[leftmargin=*]
    \item \textbf{Atmosphere:} Marble grandeur, ash‑tipped ambition, the slow rot of imperial forms, and the cloying sweetness of a city drowning in its own nostalgia.
    \item \textbf{Key Conflict:} Legitimacy versus wealth; old families, new mercenaries, and the Everflame's judgment — but beneath it all, the battle between Maelstraeus (greed) and Livaea (desire) over the souls of the powerful.
    \item \textbf{Notable Threats:} Bread riots, arena spills, ducal conflicts, inquisitorial visits, the Red List — and the Vigil's three factions, who may arrest you for debt, sedition, or simply being unnamed.
    \item \textbf{Common Rewards:} Procession permits, coin‑house credit, laurel deeds, a Golden Edict carved in precedent — and, if you are brave, a secret from the Vigil's own ledgers.
\end{itemize}

\subsection*{Prominent NPCs of Ecktoria}
\label{sec:ecktoria-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Valerius & Grand Magistrate & Preserve the forms of law while the empire crumbles & His gavel is a sundial; it only strikes at noon. He has not struck it in a year. \\
\hline
Flavia & High Priest of the Everflame & Balance the factions and keep the flame lit & She has never raised her voice; the flame speaks for her. Lately, it speaks in the voices of the dead. \\
\hline
Cassia & Duchess-Regent of the Dolmis Gate & Protect her family's trading rights & She rules from behind a silk screen; a dagger‑hole is cut at eye level. The screen was a gift from Livaea. \\
\hline
Marcus & Coin-House Factor & Survive the next credit panic & He seals deals with three ribbons: white (ask), red (now), black (never). His black ribbon is fraying. \\
\hline
Lucius & Lanista (arena master) & Own the next champion and the secrets of the hypogeum & His fighters' brands are his signature; he never repeats a mark. One brand is a spiral — the mark of the Red Vigil. \\
\hline
The Red Clerk & Censor's Hall scribe & Control which names are remembered & His inkwell is a skull; the skull belonged to a name he erased. He has forgotten his own name. \\
\hline
Vigil-Commander Vorenus & Chain-Lantern prefect & Hunt untutored magic before it wakes the Ninth & His lantern burns green. He says it shows the truth. His shadow has two heads. \\
\hline
Lady Aelia & Patrician heiress, Livaea's favourite & Outshine every rival before the next masque & She has not left the Amber Palace in seven years. She does not know the price of bread. \\
\hline
\end{longtable}

\subsection*{Names of Ecktoria}
\label{sec:ecktoria-names}

Inspired by late Latin and early Romance: strong vowels (ae, i, o), soft endings, and the weight of imperial bureaucracy. Surnames often denote a place of origin, a trade, or a notable ancestor.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Valerius & Flavia & de Urbe & \textit{the Iron} \\
\hline
Marcus & Cassia & Aquilinus & \textit{the Coin-Wise} \\
\hline
Lucius & Aelia & Honoratus & \textit{Flame-Touched} \\
\hline
Tiberius & Cornelia & Lateranus & \textit{the Unburned} \\
\hline
Gaius & Livia & Silvanus & \textit{Bread-Counter} \\
\hline
Decimus & Valeria & Portinari & \textit{Ash-Finger} \\
\hline
\end{longtable}

\subsection*{The Three Ecktorias: Faction Spectrum}
Ecktoria is not a single polity but a battlefield of three competing visions. Most citizens and factions fall somewhere between these poles, shifting allegiance as the grain price changes. Superimposed on them are the whispers of Maelstraeus and Livaea, who have no faction — only debtors.

\begin{itemize}[leftmargin=*]
    \item \textbf{Old Blood (Patrician Traditionalists)} -- The ancient senatorial families who remember when Ecktoria ruled the world. They worship precedent, hoard marble, and see mercenaries as necessary filth. Their power is in law, lineage, and the silent networks of the Censor's Hall.  
    \emph{Clocks: \textsc{Patrician Patience} [4], \textsc{Blood Debt} [6].}  
    \emph{Patron whispers:} Livaea — they marry for influence, betray for pleasure, and mistake cruelty for wit.

    \item \textbf{Black Purse (Mercenary \& Trade Leagues)} -- The new power: condotta brokers, coin‑house factors, and mercenary captains who have bought what birth could not give. They care about supply lines, not bloodlines. Their loyalty is to the next contract, and their god is the ledger.  
    \emph{Clocks: \textsc{Contract Spiral} [4], \textsc{Payday Fever} [6].}  
    \emph{Patron whispers:} Maelstraeus — they measure profit in interest, interest in suffering, and suffering in silver.

    \item \textbf{The Everflame (Church \& Inquisition)} -- The priests, inquisitors, and relic‑keepers who claim that the Eternal Flame is the only true sovereign. They offer legitimacy to whoever pays the tithe — and withdraw it from whoever forgets. Their power is in pulpits, processions, and the fear of excommunication.  
    \emph{Clocks: \textsc{Heresy Index} [4], \textsc{Sanctuary Violation} [6].}  
    \emph{Patron whispers:} Neither — the Everflame remembers, but the Vigil has learned to filter its truths.
\end{itemize}

Superimposed on these are the \textbf{Ykrul creditors} (who hold Ecktoria's debt like a leash), the \textbf{Black Banners} (who hold its streets like a lease), and the \textbf{Vigil} (who hold its secrets like a knife). A wise politician learns to balance all three — and keeps a boat ready on the Dolmis.

\subsection*{The Vigil — Three Faces of the Secret Police}
\label{sec:ecktoria-vigil}

The Vigil is not one order but three, each claiming the mantle of the old imperial secret police. They do not cooperate. They do not war. They \emph{audit} each other.

\begin{itemize}[leftmargin=*]
    \item \textbf{The Chain-Lanterns (Orthodox)} — Hunters of untutored magic. They carry brass lanterns that reveal latent talent through candle‑smoke tests. They answer to the Synod and the Everflame, but their true loyalty is to the \emph{precedent} — the law that magic must be licensed, indexed, and taxed. A Chain-Lantern raid is heralded by a single, low bell.  
    \emph{Clocks: \textsc{Candle-Smoke Sweep} [4], \textsc{Heretic Index} [6].}

    \item \textbf{The Unburnt (Patrician)} — Bodyguards and spies of the Amber Palace. They wear red sashes and carry no visible weapons. Their job is to ensure that no whisper of rebellion, no bread riot, no debt revolt ever reaches a patrician's ear. They are experts at \emph{redaction} — not killing witnesses, but making them forget.  
    \emph{Clocks: \textsc{Palace Secrets} [4], \textsc{Redacted Memory} [6].}

    \item \textbf{The Red Vigil (Maelstraeus's Own)} — Debt‑collectors and interest‑enforcers. They serve no noble, no church, no magistrate — only the ledger. They audit the wealthy for the wealthy, finding hidden obligations, unpaid tithes, and compound interest that has grown teeth. A Red Vigil agent never demands coin. They demand \emph{names}.  
    \emph{Clocks: \textsc{Interest Due} [4], \textsc{Ledger of Names} [6].}
\end{itemize}

The Vigil's three commanders meet once a month in the hypogeum of the Arena, under the gaze of statues that weep marble tears. No record is kept of these meetings. No one admits they happen. The common folk say that if you see a Vigil agent, it is already too late — but they cannot agree on \emph{which} Vigil.

\section*{Spades --- Places}
\label{sec:ecktoria-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Petition Steps} --- Milk‑stone steps where dawn petitions are pinned.
  \hfill \textit{The petitions are written in ink and hope. The steps answer in marble and silence.} `[SOCIAL]`
\item \textbf{Tally-Ford Ferryhouse} --- Rope‑drums thrum all day; the ferrymaster watches coin and banners.
  \hfill \textit{The rope‑drums have a pattern. It spells the names of those who paid late.} `[TRAVEL]`
\item \textbf{Arena Hypogeum} --- Practice sand and whispered secrets beneath the games.
  \hfill \textit{The sand is from the Tiber. It still remembers the blood of the last revolt. The Vigil meets here.} `[DEATH]`
\item \textbf{Everflame Cloister} --- Candle markets and side‑chapels where flame speaks truth.
  \hfill \textit{The flame has a voice. It is the voice of the last liar who tested it. Lately, it speaks in a whisper.} `[TRUTH]`
\item \textbf{Shatterline Bazaar} --- Between quake‑toppled columns, merchants hawk broken dreams.
  \hfill \textit{The columns were toppled by an earthquake that killed a senator. He still haunts the rubble, counting his debts.} `[TRADE]`
\item \textbf{Coin-House Floor} --- Counting ledgers under glass; factors seal credit with three ribbons.
  \hfill \textit{The ribbons are colored: white for "ask," red for "now," black for "never." The black ribbons are all frayed.} `[AUTHORITY]`
\item \textbf{Mile-Zero Obelisk} --- Surveyors' starting point; all roads measured from here.
  \hfill \textit{The obelisk has a crack. The crack is a mile long — in the other direction. The Vigil uses it to measure disappearances.} `[LAW]`
\item \textbf{Triumph Stairs} --- Processional Way switchback; marble worn smooth by ambition.
  \hfill \textit{The stairs have a name: "The Rise of the Forgotten." No one remembers why. The patricians still climb them.} `[SOCIAL]`
\item \textbf{Aqueduct Walk} --- Narrow guardpath above the city's arteries; perfect for ambush or escape.
  \hfill \textit{The water below tastes of iron. It is the city's memory of its own blood. The Vigil drinks it to stay awake.} `[STEALTH]`
\item[J] \textbf{Ducal Loggia} --- High above a packed square; a duchess watches and waits.
  \hfill \textit{The duchess has not blinked in three hours. She is counting. She has forgotten what she is counting.} `[AUTHORITY]`
\item[Q] \textbf{Censor's Hall} --- Red‑lacquered tablets record every name, every shame.
  \hfill \textit{The tablets are redacted in invisible ink. The truth is underneath. The Vigil has the reagent.} `[LAW]`
\item[K] \textbf{Grand Forum} --- Statues tarped, echoes loud; where crowds become mobs.
  \hfill \textit{The tarps were a gift from a dead emperor. The statues do not thank him. They whisper to the Vigil.} `[SOCIAL]`
\item[A] \textbf{The Amber Palace} --- Walled off from the city, where the patricians hold masques that last for weeks.
  \hfill \textit{The walls are not stone. They are debt. No one has entered without an invitation in a decade. No one has left.} `[LEGEND]`
\end{enumerate}

\section*{Hearts --- People \& Factions}
\label{sec:ecktoria-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Torchbearer Child} --- Ash‑smudged hands; knows every shortcut through the ruins.
  \hfill \textit{She has never been lost. The ruins speak to her in dust. The Vigil pays her for echoes.} `[STEALTH]`
\item \textbf{Ferrymaster} --- Knows which banners pay late; collects more than coin.
  \hfill \textit{His ledger is a net. The names are the catch. He sold a name to the Red Vigil once. He does not sleep well.} `[TRAVEL]`
\item \textbf{Coin-House Factor} --- Seals credit with three ribbons; eyes twitch at defaults.
  \hfill \textit{His eyes are the color of old silver. They see through your purse. They also see the Red Vigil in the corner.} `[AUTHORITY]`
\item \textbf{Lanista} --- Owns fighters — and their secrets; sells victory in the sand.
  \hfill \textit{His grin is a brand. The brand is on his fighters' necks. One fighter has a brand that glows — a Vigil spy.} `[COMBAT]`
\item \textbf{Decumanus-Master} --- Surveyor who "moves" stones; rewrites the city's shape.
  \hfill \textit{His rod is a finger bone. The bone belonged to the last surveyor who argued. The Vigil uses his maps.} `[LAW]`
\item \textbf{Procession Marshal} --- Permits and whistle; keeps order in ceremonial chaos.
  \hfill \textit{His whistle is for the living. His bell is for the dead. He has not used the whistle in a year.} `[SOCIAL]`
\item \textbf{Grain-Prefect} --- Allots loaves; knives in the margins of every ledger.
  \hfill \textit{His ledger has a stain. The stain is wine — or blood. The Vigil has a copy.} `[DEBT]`
\item \textbf{Censor's Clerk} --- "Misfiles" careers; knows which names to forget.
  \hfill \textit{His inkwell is a skull. The skull belonged to a name he forgot. The Vigil pays him for the ones he remembers.} `[LAW]`
\item \textbf{Veteran Standard-Bearer} --- Sells honor as a retainer; medals heavier than sword.
  \hfill \textit{His medals are from a war that never happened. The corpses are from one that did. The Vigil recruited him.} `[COMBAT]`
\item[J] \textbf{Gladiatrix} --- Darling of the crowds; owes no one but the sand.
  \hfill \textit{Her sword has no name. The sand has many. The Vigil has a file on her. It is thicker than her arm.} `[COMBAT]`
\item[Q] \textbf{Duchess-Regent} --- Behind a silk screen; speaks through cousins, rules through silence.
  \hfill \textit{The screen has a hole. The hole is for a dagger — hers. The screen was a gift from Livaea.} `[POLITICS]`
\item[K] \textbf{High Priest of Everflame} --- Smiles like law; flame judges truth and debt.
  \hfill \textit{His smile is a candle. It can warm — or burn. The Vigil has asked him to smile less.} `[TRUTH]`
\item[A] \textbf{Vigil-Commander Vorenus} --- Leads the Chain-Lanterns. Fears the Unburnt. Hates the Red Vigil.
  \hfill \textit{His lantern burns green. He says it shows the truth. His shadow has two heads — one for each faction he has betrayed.} `[AUTHORITY]`
\end{enumerate}

\section*{Clubs --- Complications/Threats}
\label{sec:ecktoria-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Blocked Bridge} --- A procession blocks the only crossing till sundown. `[TRAVEL]`
  \hfill \textit{The procession is for a saint who died hungry. The saint is not forgiving. The Vigil is watching.}
\item \textbf{Proscribed Sigil} --- A sudden edict: your mark is now forbidden. `[LAW]`
  \hfill \textit{The edict was written in the dead of night. The ink is still wet — and warm. The Vigil's seal is on the back.}
\item \textbf{Bread Riots} --- Grain barges are late; hunger hisses at the city's edges. `[SOCIAL]`
  \hfill \textit{The hunger has a face. It is the face of a baker who ran out of flour. The Vigil is looking for him.}
\item \textbf{Arena Spill} --- A riot spills into streets; shutters slam, guards draw steel. `[COMBAT]`
  \hfill \textit{The riot started over a point. The point was a sword. The Vigil started it.}
\item \textbf{Ducal Conflict} --- Rival dukes levy the same ward; your recruits vanish. `[POLITICS]`
  \hfill \textit{The recruits are not dead. They are hired. They are now your enemies — and Vigil informers.}
\item \textbf{Border Dispute} --- A re‑plat redraws boundaries; your deed becomes "disputed." `[LAW]`
  \hfill \textit{The surveyor's rod is bent. He calls it "context." The Vigil calls it leverage.}
\item \textbf{Frozen Purse} --- A coin‑house calls in notes; your purse freezes at the worst moment. `[DEBT]`
  \hfill \textit{The notes were called by a dead man. His signature is still warm. The Red Vigil is at the door.}
\item \textbf{Inquisitorial Visit} --- A relic's authenticity is questioned; penalties are swift and public. `[TRUTH]`
  \hfill \textit{The relic is a nail from a cross. It has been forged seven times. It still works. The Vigil supplied the forger.}
\item \textbf{Aftershock} --- Masonry and timetables fall; scaffolds groan with new cracks. `[DISASTER]`
  \hfill \textit{The crack has a name. It whispers it to the wind. The Vigil writes it down.}
\item[J] \textbf{Counterfeit Laurel} --- Your win is void unless you find the forger. `[SOCIAL]`
  \hfill \textit{The forger is a child. The child is a slave. The slave is a senator's son. The Vigil knows.}
\item[Q] \textbf{Church Purge} --- "Spectacles" are banned; your venue is canceled by zealots. `[DEATH]`
  \hfill \textit{The zealots are paid by a rival. The rival is in the front row. The Vigil is in the back.}
\item[K] \textbf{Rerouted Triumph} --- The crowd follows the favored path; your moment is stolen. `[SOCIAL]`
  \hfill \textit{The favored path is lined with archers. The archers are not there for the ceremony. They are Vigil.}
\item[A] \textbf{Red List} --- A secret ledger; your patron's name appears overnight. `[DEATH]`
  \hfill \textit{The list is written in a dead hand. The dead are still collecting. The Vigil holds the pen.}
\end{enumerate}

\section*{Diamonds --- Rewards/Leverage}
\label{sec:ecktoria-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Bread Chit} --- Ward sustenance for a week; street patience guaranteed. `[SOCIAL]`
  \hfill \textit{The chit is a clay token. Break it to eat — or to bribe. The Vigil accepts both.}
\item \textbf{Procession Permit} --- Move sacred flame through a crowded quarter; law walks with you. `[SOCIAL]`
  \hfill \textit{The permit is a torch. Carry it, and the city will part — or burn. The Vigil will follow.}
\item \textbf{Arena Purse} --- Settles debts and hires mouths; blood‑money in coin. `[COMBAT]`
  \hfill \textit{The purse is a helmet. The helmet belonged to a champion who lost. The Vigil has his name.}
\item \textbf{Survey Writ} --- Reset one boundary stone; redraw the city's shape. `[LAW]`
  \hfill \textit{The writ is a stone. Move it, and the map changes — on paper. The Vigil watches the paper.}
\item \textbf{Ducal Safe-Conduct} --- Four guards and a trumpet; noble favor walks ahead. `[TRAVEL]`
  \hfill \textit{The guards are paid by the duke. The trumpet is paid by you. The Vigil has a key to the trumpet.}
\item \textbf{Coin-House Credit} --- Silent and immediate; debt becomes opportunity. `[DEBT]`
  \hfill \textit{The credit is a ribbon. Cut it to spend — or to owe. The Vigil will collect the interest.}
\item \textbf{Magistrate's Postponement} --- Buy a tenday against any suit; law delayed, not denied. `[LAW]`
  \hfill \textit{The postponement is a letter. The letter is sealed with a tear. The Vigil has the other copy.}
\item \textbf{Laurel-Deed} --- Crowd favor follows you into court; glory as a shield. `[SOCIAL]`
  \hfill \textit{The deed is a leaf. It was plucked from a victor's crown. The victor is dead. The Vigil knows why.}
\item \textbf{Grain Ledger} --- Name who eats this month; hunger bends to your will. `[DEBT]`
  \hfill \textit{The ledger is a stomach. Feed it, or it will feed on you. The Vigil has the first page.}
\item[J] \textbf{Minor Title} --- A style that opens doors that bolts resist; rank as a key. `[SOCIAL]`
  \hfill \textit{The title is a word. The word is "Patrician." The doors do not ask questions. The Vigil does.}
\item[Q] \textbf{Basilica Audience} --- One private hour with the High Priest; the flame speaks truth. `[TRUTH]`
  \hfill \textit{The hour is a candle. When it burns out, so does your favor. The Vigil lit the candle.}
\item[K] \textbf{Censor's Indulgence} --- One banned thing becomes permitted (for you). `[LAW]`
  \hfill \textit{The indulgence is a silence. The silence is a scream. The Vigil has earplugs.}
\item[A] \textbf{Golden Edict} --- An exception carved in marble; precedent travels like law. `[AUTHORITY]`
  \hfill \textit{The edict is a crack. The crack in the law is now an opening. The Vigil is already inside.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:ecktoria-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (papers/favor/precedence) that change position rather than call for a roll.
\item If any A appears, echo \textbf{marble and fire} motifs — statues that watch, flames that judge, and law that remembers every name. Also echo the \textbf{Vigil's presence} — a cough in the crowd, a seal on a document you did not stamp, a mirror that shows a different face.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:ecktoria-mechanics}

\begin{itemize}
\item \textbf{Imperial Ruins:} Every district bears half‑collapsed arcades, serving as markets, courts, or shelters for mercenary companies. The Vigil knows every hidden passage.
\item \textbf{Competing Factions:} Dukes, guilds, and priests stage processions that collide at bridges and squares, each claiming legitimacy. The Vigil collects bets on the outcome.
\item \textbf{Monuments as Weapons:} Statues, obelisks, and colonnades are not only symbols — they are barricades, rallying points, and high ground when the city riots. The Vigil has a map of which statues can be toppled.
\item \textbf{Everflame Orthodoxy:} The basilicas preach unity, but each sermon is a coded signal for a political camp. The High Priest plays balance while cultivating his own following. The Vigil decodes every sermon.
\item \textbf{Black Banner Contracts:} Foreign mercenaries fill the gaps left by a shrinking citizen levy, their loyalty secured only by steady coin. The Vigil audits their pay.
\item \textbf{The Amber Palace Quarantine:} The patricians have not left the palace in a decade. No one is certain if they are prisoners or patrons. The Vigil maintains the perimeter — and delivers the wine.
\item \textbf{Local Patrons (Syncretic Saints \& Cults):}
  \begin{itemize}
    \item \textbf{St. Marcellus of the Gilded Scale (Mykkiel)} — *The Justiciar*. His order enforces unwritten laws; reading a sealed precedent binds the reader to enforce it. His shrine is the Censor's Hall. The Vigil keeps a copy of every precedent.
    \item \textbf{St. Valeria of the Serpent (Isoka)} — *The Shedding Saint*. Pilgrims who confess at her chapel forget one sin — and one kindness. Her mosaic shows a snake eating its tail. The Vigil uses her chapel to erase memories.
    \item \textbf{St. Lucia of the Undying Flame (Oath of Flame \& Light)} — *The Witness of Embers*. Any oath sworn over her shrine's flame cannot be broken; the breaker sees their own face in every wave until they drown. The Vigil swears oaths there — and never breaks them.
    \item \textbf{The Reckoner (Maelstraeus)} — *The Ledger Keeper*. A mystery cult of coin‑house factors. They offer credit at the cost of a hidden interest that compounds in blood. The Red Vigil are their debt‑collectors.
    \item \textbf{The Ember Prefect (Heretic Cult)} — A schism that believes the flame should not be contained. They meet in the arena hypogeum, chanting the *Unspoken Ninth Citation*. The Grand Magistrate has outlawed them three times. The Vigil has infiltrated them twice.
    \item \textbf{The Velvet Mask (Livaea)} — A secret society of patricians who worship pleasure as the only truth. Their rites are said to last nine days. No one has ever left a rite unchanged. The Vigil watches their parties — and reports to the Unburnt.
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross‑Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Bitter Root:} Daughters of the Cord run a hospice in the Shatterline Bazaar. They ask no payment — but they cut threads. A renegade Sister hides among the arena slaves, trying to untie her cord from a coin‑house debt. The Vigil has her file.
  \item \textbf{Malachai, the Chained Angel:} A silver coin that never spends circulates among the grain‑prefects. Take it, and your store never runs dry — but the coin will reappear in the mouth of a starving child. The Red Vigil uses it to mark debtors.
  \item \textbf{The Grey Wanderer's Grimoire:} Ecktorian psions are called “Flame‑Readers.” They see the future in the Everflame's flicker. The High Priest employs three — and has never trusted a forecast that didn't smell of smoke. The Vigil employs their apprentices.
  \item \textbf{The Threshold Compendium:} The Imperial Arena is a Class IV Poltergeist site — the collective rage of every gladiator who died for the crowd. Saikou Ira would ring a bell eight times and leave a wreath of laurel. The Vigil rings the bell nine times. The Arena answers.
  \item \textbf{The Ninth:} In any Ecktorian court, the ninth citation of precedent voids all previous. The wise stop at eight and settle. Those who cite the ninth hear a faint laughter from the marble walls — and their cases are never heard again. The Vigil cites the ninth often. The laughter does not bother them.
\end{itemize}

\subsection*{Ecktoria Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Flame:} Crumble a crust into the Everflame. Ignore the first \emph{Bread Riots} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Ribbon:} Tie a single knot in a coin‑house credit ribbon. Cancel \emph{Frozen Purse} or take --1 die on your next debt roll (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Marble:} Whisper a dead debtor's name into a statue's ear. Gain +1 Effect when collecting from their heirs this scene, but mark \emph{Obligation} to the Grand Magistrate. `[SUPERSTITION]`
  \item \textbf{Spit on the Seal:} Before signing any contract, spit on the wax seal. If it sizzles, the Vigil has a copy. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:ecktoria-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{city guards or rival factions}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{crowds or ceremonial processions}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{patrols or inquisitors}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or ceremonial opportunity}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{City authorities or noble house} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{crowd movement or procession}.
\item \textbf{Foe Appears:} A \textbf{rival faction member or city official} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{guards, mercenaries, or zealots} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes — \textbf{procession turns violent/boundary dispute erupts/patron turns}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{High Priest, Grand Magistrate, or Duchess-Regent} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically — \textbf{procession erupts/arena gate opens/monument collapses}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice — \textbf{flame speaks/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Ecktoria (Marble \& Fire):} Statues turn their heads, flames burn blue, ceremonial masks appear in windows.
\item \textbf{Ecktoria (Procession):} Crowds chant your name, banners block your path, trumpets announce your arrival.
\item \textbf{Ecktoria (Patronage):} Seals break in your hand, favors are revoked, noble houses compete for your allegiance.
\item \textbf{Ecktoria (The Vigil):} A seal appears on your documents that you did not stamp. A cough from the crowd sounds like your name. Your reflection in a polished shield has a red sash.
\end{itemize}

\subsection*{Extensions (Plug-in)}
\label{sec:ecktoria-extensions}
\begin{itemize}
  \item \textbf{Cross-Region Hooks (optional draws):}
  \begin{itemize}
    \item \emph{Aeler True Masons bid} to shore a failing aqueduct; accepting stabilizes \emph{Aftershock} but grants them soft leverage in the ward. The Vigil audits the contract.
    \item \emph{Aelinnel Green-Market broker} offers ``truth‑credit'' on court gossip; pay later in favors or names. The Vigil pays in silence.
    \item \emph{Black Banner retainer} parade demands back pay on the Triumph Stairs; crowd sympathy tilts your next parley. The Vigil arrests the ringleader.
  \end{itemize}

  \item \textbf{Faction Entanglements During Travel (use en route or on city entry):}
  \begin{itemize}
    \item If the party holds any \emph{Condotta} Diamonds, begin with \emph{Bread Riots} at +1 segment (mercenary grain trains re‑routed).
    \item Entering with an \emph{Underway Pass} (Aeler) grants +1 Position on \emph{Aqueduct Walk} scenes but ticks \emph{Ducal Conflict} once (guild jealousy). The Vigil notes your pass.
    \item Carrying \emph{Green Market} tags downgrades one \emph{Censor's Hall} action but flags \emph{Inquisitorial Visit} at the next temple. The Vigil compares tags.
  \end{itemize}

  \item \textbf{Trade Good Consequences (pick 1 when a haul sells):}
  \begin{itemize}
    \item \emph{Arena Steel} $\rightarrow$ street duels spike $\rightarrow$ start \emph{Arena Spill} at 1. The Vigil collects the weapons.
    \item \emph{Relic Fragments} $\rightarrow$ basilica prestige swings $\rightarrow$ tick \emph{Church Purge} or grant \emph{Basilica Audience}. The Vigil inspects the relic.
    \item \emph{Survey Rights} $\rightarrow$ boundary re‑plat $\rightarrow$ downgrade one \emph{Border Dispute} \emph{or} upgrade a rival's. The Vigil updates the map.
  \end{itemize}

  \item \textbf{Diplomatic Favor (minor currency):}
  \begin{itemize}
    \item Earn by resolving a public scene without bloodshed (Forum, Procession, Coin‑House).
    \item Spend 1 to reduce DV of a permit/audience by 1, or to ignore \emph{Blocked Bridge} for one crossing.
    \item Holding 2+ makes you ``interesting'': start a \emph{Red List} [4] that advances on snubbed patrons. The Vigil adds your name.
  \end{itemize}

  \item \textbf{Reputation Echoes (soft power):}
  \begin{itemize}
    \item \emph{Crowd-Loved (Laurel-Deed):} first time per session you face a mob, set Position one step safer. The Vigil watches how the mob loves you.
    \item \emph{Coin-House Trusted (Credit):} ignore the first \emph{Frozen Purse}; next time, it returns at +1 severity. The Vigil audits your credit.
    \item \emph{Clerically Vouched (Everflame):} downgrade one \emph{Inquisitorial Visit}; if abused, promote \emph{Church Purge}. The Vigil has a copy of your voucher.
  \end{itemize}
\end{itemize}

\subsection*{Monuments as Mechanics (opulence with teeth)}
\begin{itemize}
  \item \textbf{Processional Geometry:} When acting on the \emph{Triumph Stairs}, choose: +1 Effect \emph{or} --1 Heat; on a miss, \emph{Rerouted Triumph} advances. The Vigil stands at the top of the stairs, watching.
  \item \textbf{Aqueduct Leverage:} Securing an aqueduct span grants a one‑scene \emph{high ground} tag in any linked ward; if collateral hits, tick \emph{Aftershock}. The Vigil knows where the aqueduct leaks.
  \item \textbf{Forum Acoustics:} Rally speeches at the \emph{Grand Forum} grant an extra assist die if at least 20 witnesses; failure spawns \emph{Bread Riots} at +1. The Vigil counts the witnesses.
\end{itemize}

\subsection*{Plug-in SB Conversions (Ecktoria flavor)}
\begin{itemize}
  \item Any 1 SB may become \emph{Crowd Drift}: shift the scene's terrain in favor of or against you (bridges clog, stairs clear).
  \item Any 2 SB may become \emph{Seal Revoked}: void one paper (Permit/Chit/Indulgence); to avert, sacrifice \emph{Diplomatic Favor}.
  \item On 3+ SB offer \emph{Patron Bargain}: accept a visible mark (ribbon, laurel brand) to halve consequences; start a \emph{Patron Debt} [6].
\end{itemize}

% ====== ANCIENT CONSPIRACIES, CULTS, & SCHISMS ======
\subsection*{Hidden Orders \& Ancient Conspiracies (d12 hooks)}
\label{sec:ecktoria-conspiracies}
\begin{enumerate}
  \item \textbf{The Ash Senate} -- a ledger of ``extinguished'' patrician lines still voting by proxy through funerary guilds. The Vigil has the key.
  \item \textbf{Surveyors of the Ninth Mile} -- decumanus‑masters who can \emph{move} boundaries with ritual math; a lost rod sets precedent across the city. The Vigil owns the rod.
  \item \textbf{Lantern Tetrarchy} -- four basilica beadles who decide which miracles are counted; the fifth lamp was broken in a purge. The Vigil lit the purge.
  \item \textbf{The Red Marble Compact} -- coin‑house founders who tithe 1 in 100 coins to a secret indemnity fund used to flip coups. The Red Vigil administers the fund.
  \item \textbf{Archivum Umbrae} -- censor clerks maintaining a \emph{shadow index} of names that must not stand trial (yet). The Vigil reads the index nightly.
  \item \textbf{Peregrine Brotherhood} -- ferrymasters trading in sealed confessions; blackmail that floats. The Vigil buys the confessions.
  \item \textbf{The Laurel of Dust} -- a gladiators' mutual that chooses which champions live long enough to hold office. The Vigil chooses the champions.
  \item \textbf{Aqueduct Collegium Obscura} -- engineers who keep a second set of sluice keys; droughts become elections. The Vigil has the third set.
  \item \textbf{Order of the Broken Obelisk} -- survey‑cult that believes Mile‑Zero is \emph{wrong} and can re‑found the city with one stone. The Vigil has the stone.
  \item \textbf{The Funeral Republic} -- undertakers running a parallel courier state through catacombs and wakes. The Vigil uses their couriers.
  \item \textbf{The Ember Vault} -- a sealed reliquary under the Arena; victors' sweat was bottled for augury. The Vigil broke the seal.
  \item \textbf{Triumphal Maskers} -- procession costumers who encode heretical theses into robe pleats and torch patterns. The Vigil decodes the theses — and sometimes agrees.
\end{enumerate}

\subsection*{Cults \& Mystery Rites (d10)}
\label{sec:ecktoria-cults}
\begin{enumerate}
  \item \textbf{Children of the Phoenix} (\emph{rebirth by fire}): ash‑marks grant +1 die vs.\ fear in crowds; mark Fatigue if you hesitate near open flame.
  \item \textbf{Verdant Laurel} (\emph{blood to marble, sap to coin}): once/session turn 1 Harm into 1 Coin‑House Credit; start a \emph{Green Debt} [4].
  \item \textbf{Mask of the Thirteenth} (\emph{forbidden procession step}): walk through a riot unmolested; next parley with clergy begins one step worse.
  \item \textbf{Brotherhood of the River Tongue} (\emph{truth ferried at night}): ask the Dolmis one yes/no; a ferryman shows up later to collect a name.
  \item \textbf{Candlemakers of Ever} (\emph{flame remembers lies}): blue flame points to falsehood; the liar dreams of you for three nights (Exposure +1).
  \item \textbf{Sable Standard} (\emph{war saint cult}): your banner ignores morale checks once; start a \emph{Blood-Tithe} [4] owed to the saint's chapel.
  \item \textbf{The Ninth Bell} (\emph{hour that is not kept}): skip a queue or sentence; next edict you need arrives late (Delay SB free once).
  \item \textbf{Guild of Red Accountants} (\emph{sin as ledger}): convert 2 SB into a \emph{Magistrate's Postponement}; add \emph{Karmic Debt} [6].
  \item \textbf{The Marble Eaters} (\emph{mortification sect}): bite stone, gain +1 Effect resisting crowd crush; take Harm 1 (Stress) if you boast.
  \item \textbf{Torch of Saint Vessa} (\emph{saint of necessary arson}): sanctify a fire to purge a tag (Plague/Corruption); City Heat +2.
\end{enumerate}

\subsection*{Religious Schisms \& Trials}
\label{sec:ecktoria-schisms}

\paragraph{The Everflame Divide (choose a live schism)}
\begin{enumerate}
  \item \textbf{Orthodox Basilica vs.\ Ember Reformers:} Is miracle \emph{counted} by law or \emph{kindled} by need?
  \item \textbf{Laurel Theologians vs.\ Street Saints:} Does glory flow \emph{down} from triumphs or \emph{up} from bread‑lines?
  \item \textbf{Iconoclast Ward vs.\ Reliquary Keepers:} Is the flame purer without images?
\end{enumerate}

\paragraph{Schism Mechanics (table-facing)}
\begin{itemize}
  \item \textbf{Heresy Index} [0--6]: ticks when players publicize a stance, defy clergy, or wield cult rites in view. At 3: inquisitorial audits become common; at 6: \emph{Church Purge} (Q) auto‑triggers once/session until lowered.
  \item \textbf{Sanctuary / Anathema Tags:} locations acquire tags during arc scenes. \emph{Sanctuary:} first harm in scene becomes Fatigue; \emph{Anathema:} first social miss spawns ``Proscribed Sigil.''
  \item \textbf{Trial by Fire (mini‑procedure, 3 beats):}
  \begin{enumerate}
    \item \emph{Assembly:} set the forum/stair as arena; state thesis in one sentence (players or NPCs).
    \item \emph{Witness:} three proofs (relic, miracle, precedent). Each proof: test Social/Arcana/Resolve vs.\ crowd DV; on success, reduce Index by 1 or gain \emph{Basilica Audience}.
    \item \emph{Judgment:} roll Crowd Mood (1d6): 1--2 mobs, 3--4 silent acceptance, 5--6 laurel toss (gain \emph{Laurel-Deed}). If Index \(\geq 4\), --1 to the roll.
  \end{enumerate}
\end{itemize}

\subsection*{Ecktoria -- Marble and Fire}
\textbf{Starting Location:} The Coin‑House floor, marble cracked with soot lines, where votive candles burn beside melted icons.
\begin{quote}
\textit{``In Ecktoria, even faith must balance its ledgers, and the gods audit in ash.''}
\end{quote}
\textbf{In...} the Coin‑House floor, a factor with smoke‑singed ledgers offers a purse. ``The Lame King moves,'' she says. ``And the Basilica burns with him. The Amber Palace has not answered a summons in a month. The Vigil is asking questions. I suggest you find the answers before they find you.''

\paragraph{Regional Motifs} Coin‑oaths, basilica relics, marble acoustics, augury smoke, Vigil seals.
\paragraph{Whispers to Offer} ``Bid against the city,'' ``Pocket the relic shard,'' ``Name an innocent debtor,'' ``The Vigil's informer sits in the corner.''
\paragraph{External Clocks} Threat: \textit{Procession of Ash [6]}; Community: \textit{Credit Panic [4]}; Anchor: \textit{Basilica Choir [4]}; Vigil: \textit{Red List [6]}.
\paragraph{Complications} Smoke speaks truths; coins stick to skin; a reliquary runs hot; a seal appears on your documents that you did not stamp.
\paragraph{Anchors at Risk} A choir novice; a patron account; a Vigil informer who owes you a favour.
\paragraph{Escalation Beat} The Basilica tolls a note only one PC hears. The Vigil hears it too.
\paragraph{Relief} Return the shard; pay a city‑debt in blood or \Boons{}; route the procession elsewhere; give the Vigil a name.

\subsection*{Boss Seeds (Cults \& Conspiracies)}
\begin{itemize}
  \item \textbf{The Ember Prefect} (schismatic inquisitor): \emph{Phase} [8] Citation $\rightarrow$ Censure $\rightarrow$ Auto‑da‑fé $\rightarrow$ Purge. Lair: Everflame Cloister. Counters: relic provenance, crowd‑favor, rival sermon. \emph{Vigil twist:} The Prefect was once a Vigil informer. The Vigil knows where he hides.
  \item \textbf{The Surveyor-Primus} (geomantic usurper): \emph{Phase} [6] Re‑plat $\rightarrow$ Dispossess $\rightarrow$ Erase. Lair: Mile‑Zero. Counters: Aeler measure‑rods, witness line, broken rod. \emph{Vigil twist:} The Primus has a Red Vigil agent as a protégé.
  \item \textbf{Phoenix of the Arena} (miracle or hoax?): \emph{Phase} [10] Hype $\rightarrow$ Immolation $\rightarrow$ ``Rebirth'' $\rightarrow$ Riot. Lair: Imperial Arena. Counters: doused ash, Mask of the Thirteenth, Everflame test. \emph{Vigil twist:} The Phoenix is a Vigil experiment. The Unburnt want to contain it. The Red Vigil wants to sell tickets.
  \item \textbf{The Velvet Mask} (Livaea's cult): \emph{Phase} [8] Invitation $\rightarrow$ Masque $\rightarrow$ Debt of Pleasure. Lair: The Amber Palace. Counters: an unbroken oath, a mirror with no reflection, a name spoken without desire. \emph{Vigil twist:} The Vigil's Unburnt faction attends every masque. They are not there to stop it.
\end{itemize}

\subsection*{Diamonds: Forbidden Leverage (add to \S\ref{sec:ecktoria-rewards})}
\begin{enumerate}
\setcounter{enumi}{13}
  \item \textbf{Shadow Index Leaf} -- remove one name from the Red List for a week; add \emph{Countercoup} +1.
  \item \textbf{Ember Writ} -- sanction one cult rite as ``orthodox'' for a scene.
  \item \textbf{Surveyor's Ninth Rod} -- redraw one ward line unnoticed (until dawn).
  \item \textbf{Vigil Seal (Red)} -- For one scene, the Red Vigil treats you as a contractor. Interest will be collected later.
\end{enumerate}

\subsection*{Cross-Contamination (fast ties)}
\begin{itemize}
  \item \textbf{Aeler True Masons:} backing an iconoclast audit grants shoring crews; bell‑codes now gate temple entries. The Vigil has the bell‑codes.
  \item \textbf{Aelinnel Thorn Courts:} hawthorn chaplains host ``mirror sermons''; a basilica gains \emph{Sanctuary} while a rival becomes \emph{Anathema}. The Vigil watches both.
  \item \textbf{Black Banners:} war‑chaplains monetize indulgences; clearing \emph{Payday} risks +1 \emph{Heresy Index}. The Vigil takes a cut.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Ecktoria's decline is not collapse, but transformation. Wealth bleeds into patronage webs: a senator's indulgence, a censor's indulgence, a noble's purse. The city's great houses keep mercenary captains on retainer, while priests sell ritual processions as public theater and private favor. And everywhere, the Vigil watches — not to stop the rot, but to take notes.

Yet beneath the decay, two demons whisper. \textbf{Maelstraeus} teaches that every contract is a weapon, every debt a leash. \textbf{Livaea} teaches that every desire is a chain, every pleasure a payment. The patricians who once ruled the world now rule only the Amber Palace, and the Amber Palace is a gilded cage. They do not see the bread riots. They do not hear the Vigil's footsteps. They dance. They drink. They forget that the flame remembers everything.

\textbf{For the GM:}  
Patronage should function as both opportunity and trap. Any gift (a purse, permit, or charter) binds the receiver to service, but patrons can die, fall from favor, or suddenly proscribe what they once allowed. The Vigil offers a different kind of patronage: \emph{protection} in exchange for \emph{information}. A party that pays the Vigil in secrets may find doors opening — and other doors closing behind them. To emphasize the decadence and dread:
\begin{itemize}
\item Tie rewards to visible symbols (seals, ribbons, laurel crowns, Vigil marks) that can be stolen, forged, or voided.
\item Let rival patrons issue conflicting "safe‑conducts," forcing players to choose whose favor matters more.
\item Use processions, funerals, or trials as arenas for social contests, where patronage debts surface and shift allegiances.
\item Introduce the \textbf{Vigil's Faction Clock [6]} to track which Vigil faction (Chain-Lanterns, Unburnt, Red Vigil) has the most leverage over the party. When it fills, that faction makes a demand.
\end{itemize}
Patronage in Ecktoria is not mere background — it is the engine of politics, coin, and survival. And beneath it all, Maelstraeus and Livaea are writing a contract. The city is the collateral. The poor are the interest. The patricians are the signature — and they signed in blood long ago.
\end{tcolorbox}

% Index entries
\index{Ecktoria|see{Marble and Fire}}
\index{Theme generator}
\index{Marble and Fire generator}
\index{Places!Ecktoria}
\index{People!Ecktoria}
\index{Complications!Ecktoria}
\index{Rewards!Ecktoria}
\index{Church of the Flame!Ecktoria}
\index{Everflame}
\index{Imperial Arena}
\index{Grand Forum}
\index{Censor's Hall}
\index{Coin-house}
\index{Duchess-Regent}
\index{High Priest of the Everflame}
\index{Grand Magistrate}
\index{Gladiatrix}
\index{Lanista}
\index{Decumanus-master}
\index{Triumphal procession}
\index{Golden Edict}
\index{Laurel-deed}
\index{Shatterline bazaar}
\index{Surveyors' Mile-Zero}
\index{Old Blood}
\index{Black Purse}
\index{Everflame faction}
\index{Saint Marcellus}
\index{Saint Valeria}
\index{Saint Lucia}
\index{The Reckoner}
\index{Ember Prefect}
\index{Order of the Unburdened}
\index{Amber Palace}
\index{Vigil (secret police)}
\index{Chain-Lanterns}
\index{Unburnt (Vigil faction)}
\index{Red Vigil}
\index{Maelstraeus (demon)}
\index{Livaea (demon)}
