\section{The Himadri Foothills --- ``The Roof of the World''}
\label{chap:himadri}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The mountain does not negotiate. It waits. The ninth breath is a toll, not a gift.}

\begin{quote}
  \textbf{Khardung, Vent-Prior of the High Places (Elite):}
  ``The mountain does not speak in words. It speaks in the silence after a bell, the weight of a frozen lung, the taste of iron on the tongue at dusk. We Aeler do not conquer stone. We listen to it. And we remember what it forgets.''
\end{quote}

\begin{quote}
  \textbf{Mine-Breath (Commoner):}
  ``I carry a bell that counts my breaths. Eight is safe. Nine is the mountain's. My grandfather counted to nine once. The mountain took his shadow. He lived another thirty years, but he never cast a shadow again. The High Folk say the shadow still walks the passes, counting.''
\end{quote}

\textbf{Genre / Mood:} High-altitude survival horror, animist dread, isolation, the weight of ancient oaths.

\textbf{Starting Location:} A wind-scoured pass where prayer flags made of frozen leather snap like bones. A dwarf in a patched iron coat tends a brass brazier. Beside her, a human in a grey wool cloak --- face hidden by a veil of yak hair --- holds a bell that does not ring. The dwarf speaks first: ``The High Folk say the demon stirs in the deep. The passes are closing early. You have three breaths to declare your purpose, or the mountain will decide for you.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Himadri foothills are the highest inhabited place in the known world. Here, the Aelerian mountains scrape the sky, and the air is so thin that each breath is a measured gift. The native Aeler of these peaks --- called the High Aeler or the \emph{Karak-Aeler} (``Stone-Breath People'') --- are not the mercantile dwarves of the lowland holds. They are severe, cold, and ancient, their faces carved by wind and frost, their voices rarely raised above a whisper. They do not trade in breath-promises for profit; they trade in survival. A Karak-Aeler's word is iron not because of law, but because the mountain has no patience for lies.

The humans of the Himadri are equally remote: a scattering of monastic villages, goat-herding clans, and hidden fortress-temples that have never paid tribute to any king. Outsiders call them the High Folk, or simply the \emph{Mi-tsang} --- the Hidden People. They speak a language of tones and silences, and they venerate spirits that live in glaciers, hot springs, and the bones of extinct beasts. Their lamas (called \emph{Lopön}) keep scrolls that predate the Utaran Empire, written in a script that no Aeler can read. They do not trust the lowlands, and the lowlands have learned not to cross their passes without an invitation.

The Karak-Aeler and the High Folk share an uneasy peace, brokered centuries ago when the first iron was traded for the secret of the high passes. Today, the dwarves mine the frozen mountains for ore and rare earths, while the humans tend the yak herds and the shrines that keep the mountain spirits from waking. The Covenant has a small mission in the highest valley, but its priests are always on the edge of frostbite and despair. The demon sealed in the Ninth Seal Mine --- a fragment of Varash the Unmaker --- is not a legend. It is a fact. The seal is cracking. And the mountain is listening.

\vspace{2mm}
\textbf{What you notice first:} The silence. Not the absence of sound, but the weight of it --- as if the mountain itself is holding its breath. The way every Aeler touches a brass bell at their belt before speaking of distances. The way the High Folk never look directly at the peaks, only at the ground, as if the sky holds something they are not meant to see.

\textbf{The one rule that kills newcomers:} Never count your breaths aloud in a mine or a high pass. The mountain listens, and if it hears you reach the ninth, it will take one --- a breath, a memory, a shadow. The High Folk say that the shadow then walks the passes, counting.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Peak}

Power here is measured in altitude, in the number of passes you know, and in the respect the mountain shows you. The Karak-Aeler maintain the iron mines and the toll-bridges, but they do not rule the High Folk --- they coexist with them. The High Folk's lamas negotiate on behalf of the mountain spirits, and the dwarves have learned to listen. The Covenant's seal on the Ninth Seal Mine is maintained by a single, half-mad priest who has not seen the sun in forty years. The demon whispers promises of warmth, of gold, of the lowlands' green fields. The miners cover their ears. The High Folk leave offerings of frozen butter tea at the seal's entrance. So far, it has held.

\subsection*{Prominent NPCs of the Himadri Foothills}
\label{sec:himadri-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Khardung Stone-voiced & Vent-Prior of the High Places & Keep the mines open and the demon sealed & He has not slept in a heated room in forty years. He says warmth makes the mountain forget him. \\
\hline
Lopön Tenzin & High Folk lama of the Ice Shrine & Interpret the will of the mountain spirits & He has not spoken a word of Low Utaran in thirty years. He communicates through a novice who signs. \\
\hline
Mine-Captain Dolkhar & Karak-Aeler engineer & Extract enough ore to satisfy the mountain's demand & She lost three fingers to frostbite; she replaced them with iron picks. She never removes her gloves. \\
\hline
The Seal-Keeper (Nameless) & Covenant priest of the Ninth Seal & Keep the demon from waking, one day at a time & He has forgotten his own name. The seal remembers it for him. His eyes are rust-coloured. \\
\hline
Yak-Mother Pema & High Folk herder, keeper of the pass offerings & Ensure the mountain spirits do not close the high route & She knows every hidden spring and every avalanche chute. She will not guide outsiders --- only sell them directions. \\
\hline
\end{longtable}

\subsection*{Names of the Himadri Foothills}
\label{sec:himadri-names}

Karak-Aeler names are short, harsh, with guttural stops and iron-hard consonants. High Folk names are softer, with flowing vowels and honorific prefixes.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{Karak-Aeler (Male)} & \textbf{Karak-Aeler (Female)} & \textbf{High Folk (Both)} & \textbf{Epithets} \\
\hline
Khardung & Dolkhar & Tenzin & \textit{Iron-Beard} \\
\hline
Borin & Thrain & Pema & \textit{Stone-Breath} \\
\hline
Rurik & Korra & Lhamo & \textit{the Unshadowed} \\
\hline
Halik & Svar & Namgyal & \textit{Bell-Breaker} \\
\hline
\end{longtable}

\paragraph*{(Mine/Shrine/Pass)} Iron mine with a carved face that weeps rust; a High Folk shrine of stacked stones and frozen prayer flags; a pass where the air is so thin that even the crows turn back.

\section*{Spades --- Places}
\label{sec:himadri-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Iron Mine of the Crying Face} --- A deep shaft with a carved face above the entrance.
  \hfill \textit{The face weeps rust. The tears are iron oxide. The mountain is mourning its dead.} \texttt{[INDUSTRIAL]}
\item \textbf{The Ninth Seal Mine (Sealed)} --- A tunnel mouth sealed with Covenant clay and iron bands.
  \hfill \textit{The seal is warm. It glows faintly at night. The demon whispers through the cracks in a language that sounds like butter melting.} \texttt{[DEATH]} \texttt{[MAGIC]}
\item \textbf{Silent Pass} --- A high mountain pass where no bird flies.
  \hfill \textit{The air is too thin for sound. Arguments must be written in the snow. The snow does not lie.} \texttt{[WEATHER]}
\item \textbf{Bridge of Nine Echoes (Karak-Aeler)} --- A rope bridge over a frozen chasm.
  \hfill \textit{Each step echoes nine times. The ninth echo is always a whisper --- the name of the next person to fall. Do not answer.} \texttt{[DEATH]} \texttt{[FEAR]}
\item \textbf{Ice Shrine} --- A cave temple where the High Folk leave offerings of butter tea and frozen barley.
  \hfill \textit{The ice never melts, even when a fire is lit inside. The lamas say the mountain drinks the heat.} \texttt{[RITUAL]}
\item \textbf{Yak-Mother's Camp} --- A seasonal encampment of black wool tents, smelling of smoke and sour milk.
  \hfill \textit{The yaks have names. They answer to them. Outsiders who guess a yak's name correctly are offered tea. Those who guess wrong are offered the cold.} \texttt{[COMMUNITY]}
\item \textbf{The Frozen Record (Vent-Prior's Archive)} --- A cave where breath-promises are carved into slabs of permafrost.
  \hfill \textit{The ice preserves the contracts. If a promise is broken, the slab cracks. The sound echoes through the mines for days.} \texttt{[AUTHORITY]}
\item \textbf{High Folk Watchtower} --- A stone spire on a ridge, manned by a silent figure in a grey cloak.
  \hfill \textit{The watcher never sleeps. They signal with mirrors --- one flash for safe, two for storm, three for demon-walk.} \texttt{[WATCH]}
\item \textbf{The Ninth Breath Shrine} --- A small cairn where miners leave offerings of their own shadow-traces.
  \hfill \textit{The offerings are not coin --- they are memories written on strips of leather. The leather never rots. The memories are still warm.} \texttt{[RITUAL]}
\item[J] \textbf{Abandoned Mine of the First Seal} --- A collapsed tunnel where the demon first manifested.
  \hfill \textit{The walls are coated in green glass --- like the Kalrantha amulet. The glass sings at midnight. Those who listen too long forget their names.} \texttt{[MAGIC]} \texttt{[DEATH]}
\item[Q] \textbf{The Lopön's Hermitage} --- A tiny stone hut perched on a cliff, accessible only by a rope ladder.
  \hfill \textit{The lama has not descended in nine years. His food is hauled up by a single, silent novice. The novice never speaks of what the lama sees.} \texttt{[MYSTERY]}
\item[K] \textbf{Karak-Aeler Forge (High)} --- A smithy built into a volcanic vent, heated by the mountain's own warmth.
  \hfill \textit{The forge has no chimney. The smoke is swallowed by the stone. The smiths say the mountain eats their impurities.} \texttt{[CRAFT]}
\item[A] \textbf{The First Keystone (Myth)} --- A stone that never cracks, said to lie at the deepest point of the oldest mine.
  \hfill \textit{It is the mountain's heart. Touch it, and the mountain will ask a question. Answer poorly, and the passes close forever. Answer well, and it may ask another.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Dwarf/Lama/Herder)} Karak-Aeler mine-captain with a bell and a level; High Folk lama with a veil of yak hair; yak-mother with a staff of frozen juniper.

\section*{Hearts --- People \& Factions}
\label{sec:himadri-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Mine-Captain (Karak-Aeler)} --- A dwarf with a brass bell and a level. She never counts her breaths aloud.
  \hfill \textit{She has never lost a crew. She has also never smiled. The miners say her bell rings only for the dead.} \texttt{[AUTHORITY]}
\item \textbf{Lopön (High Folk Lama)} --- A figure in grey wool, face hidden by a veil. They communicate through a novice who signs.
  \hfill \textit{The lama's hands are stained black with iron from the offerings. They never touch the ground.} \texttt{[SPIRIT]}
\item \textbf{Yak-Mother Pema} --- A weathered woman with a bone needle in her hair, always surrounded by yaks.
  \hfill \textit{She will not guide you --- but she will sell you directions for a story. The story must be true.} \texttt{[COMMUNITY]}
\item \textbf{The Seal-Keeper (Nameless)} --- A Covenant priest in rust-stained robes, sitting by the Ninth Seal.
  \hfill \textit{He has not eaten solid food in forty years. The seal's glow sustains him. He has forgotten his own name. The seal remembers it for him.} \texttt{[JUSTICE]}
\item \textbf{Vent-Prior Khardung} --- A dwarf with a bell around his neck, his face cracked by frost.
  \hfill \textit{The bell rings when he lies. It rings often. He is always lying about the ore quotas. The miners know. They work anyway.} \texttt{[AUTHORITY]}
\item \textbf{High Folk Watcher} --- A silent figure in a watchtower, holding a brass mirror.
  \hfill \textit{They have not spoken in nine years. They communicate with mirror flashes. One flash: safe. Two: storm. Three: demon-walk. Four has never been seen.} \texttt{[SCOUT]}
\item \textbf{Iron-Bound Miner} --- A dwarf who traded a memory to the mountain for a higher quota. She can no longer recall her mother's face.
  \hfill \textit{She works in the deepest galleries. The other miners leave offerings of silence when she passes.} \texttt{[BURDEN]}
\item \textbf{High Folk Novice} --- A young woman who speaks for the silent lama. Her hands are always wrapped in wool.
  \hfill \textit{She has never spoken a word of her own. The lama speaks through her. She does not know if she is a vessel or a person.} \texttt{[MYSTIC]}
\item \textbf{Frozen Caravan Master} --- A Dhaharan trader who braves the high passes for rare earths.
  \hfill \textit{He has lost three fingers to frostbite. He keeps them in a jar as a warning. The jar is on the ninth shelf.} \texttt{[TRADE]}
\item[J] \textbf{Ghost of the Landslide} --- A figure in grey, seen only at dusk. It points at the pass.
  \hfill \textit{It is pointing at a safe route --- or a trap. No one has survived to find out. The ghost does not speak. It does not need to.} \texttt{[DEATH]}
\item[Q] \textbf{Demon's Whisper} --- A voice heard in the Ninth Seal Mine. It offers deals.
  \hfill \textit{``A year of warmth for a single memory. A century of life for a death. The seal is cold. I am patient.'' The voice sounds like someone you loved.} \texttt{[MAGIC]} \texttt{[FEAR]}
\item[K] \textbf{The First Miner (Legend)} --- A skeleton holding a tuning fork, sitting at the First Keystone.
  \hfill \textit{If you take the tuning fork, you become the mountain's new voice. You will never leave the depths. The miners will hear you sing. They will not thank you.} \texttt{[LEGEND]}
\item[A] \textbf{Ice-Lama (Myth)} --- A High Saint who froze to death in meditation. Her body did not decay.
  \hfill \textit{She sits in a cave of blue ice. If you bring her a true regret, she will warm your hands for a moment. The warmth lasts a lifetime.} \texttt{[SAINT]}
\end{enumerate}

\paragraph*{(Cold/Seal/Spirit)} A sudden drop in temperature --- all actions Desperate; the Ninth Seal cracks --- the demon whispers promises; a mountain spirit demands an offering of silence.

\section*{Clubs --- Complications/Threats}
\label{sec:himadri-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Whiteout} --- A sudden blizzard erases all tracks. Navigation is impossible. \texttt{[WEATHER]}
  \hfill \textit{The snow tastes of iron. Someone is bleeding nearby. The wind sounds like a name being called. Your name.}
\item \textbf{Landslide} --- The pass is blocked. The alternative route is three days longer. \texttt{[TRAVEL]}
  \hfill \textit{The landslide was deliberate --- a rival caravan master wants to delay you. His camels are ahead. His guilt is written in the mud.}
\item \textbf{Ninth Echo Heard} --- Someone on the Bridge of Nine Echoes counts the echoes. The ninth echo speaks a name. \texttt{[DEATH]} \texttt{[FEAR]}
  \hfill \textit{The name is the speaker's. They faint. The bridge remembers. The name is now a price. The bridge will collect.}
\item \textbf{Breath Shortage} --- At high altitude, the party struggles. Mark Fatigue 1 per scene. \texttt{[WEATHER]} \texttt{[SURVIVAL]}
  \hfill \textit{Acclimatisation requires a rest day. The rest day will cost a price. The price is a breath-promise. The mountain is watching.}
\item \textbf{Seal Cracking} --- The Ninth Seal groans. The demon's whisper grows louder. \texttt{[DEATH]} \texttt{[MAGIC]}
  \hfill \textit{A character hears their own voice offering a deal. The price is a memory of a warm place. Refuse, and the cold deepens.}
\item \textbf{Yak Stampede} --- A herd of startled yaks thunders through the camp. \texttt{[DISASTER]}
  \hfill \textit{They are not angry. They are fleeing something worse. The mist behind them has a shape. The shape is counting.}
\item \textbf{Covenant Audit} --- The Seal-Keeper demands to inspect the party's cargo. \texttt{[JUSTICE]} \texttt{[AUTHORITY]}
  \hfill \textit{The cargo includes a relic from the Red Desert --- a piece of green glass. It is humming. The Seal-Keeper's eyes are weeping rust.}
\item \textbf{High Folk Taboo Broken} --- A party member speaks a forbidden name or forgets to offer butter tea at a shrine. \texttt{[SPIRIT]}
  \hfill \textit{The mountain spirits are offended. The next pass they cross will be twice as long. The snow will fall upward.}
\item \textbf{Demon-Touched Miner} --- A miner who made a bargain with the demon escapes from the deep galleries. \texttt{[COMBAT]}
  \hfill \textit{Their eyes are green glass. Their shadow has horns. They do not attack --- they ask: ``Where is the ninth bone?''}
\item[J] \textbf{Vent-Prior's Summons} --- Khardung demands the party attend him. He offers a deal: find a saboteur in exchange for warm shelter. \texttt{[AUTHORITY]} \texttt{[POLITICS]}
  \hfill \textit{The saboteur is his own clerk. The clerk is his daughter. She is trying to open the Ninth Seal. She wants to free the demon. She thinks it will warm the mines.}
\item[Q] \textbf{The Silent Pass's Cold (Supernatural)} --- The temperature drops so low that metal sticks to skin. \texttt{[WEATHER]} \texttt{[MAGIC]}
  \hfill \textit{The cold is a spirit. It wants a breath-promise. It will accept a story. The story must be about fire. It has never heard a good one.}
\item[K] \textbf{Demon Wakes} --- The Ninth Seal cracks open. A shadow escapes. \texttt{[DEATH]} \texttt{[MAGIC]}
  \hfill \textit{It possesses the nearest High Folk novice. The novice's eyes turn to green glass. She speaks in the lama's voice: ``The mountain is hungry. Feed it.''}
\item[A] \textbf{First Keystone Quakes} --- The mountain's heart trembles. The passes close. The mines collapse. The demon is free. \texttt{[APOCALYPSE]}
  \hfill \textit{The party has one day to re-seal the demon or flee. The Vent-Prior is weeping. The Seal-Keeper is dead. The mountain is counting.}
\end{enumerate}

\paragraph*{(Bell/Ingot/Prayer)} Aeler breath-counter (silver bell, wear to count breaths safely); iron ingot stamped with a mountain rune; a High Folk prayer flag that never tears.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:himadri-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Aeler Breath-Counter (Silver Bell)} --- Wear it to count your breaths safely in high altitude. The bell rings when you reach eight. \texttt{[TRAVEL]} \texttt{[SURVIVAL]}
  \hfill \textit{Stop at eight. The ninth breath is the mountain's. The bell also rings when a demon is near. The tone is flat.}
\item \textbf{Iron Ingot (Mountain-Stamped)} --- A bar of high-grade iron, stamped with a rune that means ``witness.'' Worth 1 Measure. \texttt{[TRADE]}
  \hfill \textit{The rune glows when a lie is spoken nearby. It does not judge the lie --- it merely records.}
\item \textbf{High Folk Prayer Flag} --- A square of undyed wool, marked with a symbol of a closed eye. \texttt{[PROTECTION]}
  \hfill \textit{Hang it at your camp, and the mountain spirits will not trouble you for one night. The flag must be burned at dawn. The smoke carries your thanks.}
\item \textbf{Seal-Keeper's Blessing (Clay Amulet)} --- A lump of clay from the Ninth Seal's entrance. \texttt{[JUSTICE]} \texttt{[MAGIC]}
  \hfill \textit{Press it to a lock or a threshold, and the demon's attention is drawn elsewhere for one hour. The amulet crumbles after use. The demon will remember.}
\item \textbf{Yak-Mother's Route (Leather Map)} --- A scrap of leather marked with charcoal lines. \texttt{[TRAVEL]}
  \hfill \textit{The map shows a safe route through the high passes --- but only if you follow it exactly. If you stray, the map rewrites itself. The new route leads to a crevasse.}
\item \textbf{Lopön's Silence (Wool Cord)} --- A braided cord that the lama wore around his wrist. \texttt{[SOCIAL]} \texttt{[SPIRIT]}
  \hfill \textit{Tie it around your own wrist. For one scene, you are invisible to mountain spirits. You also cannot speak. The silence is the price.}
\item \textbf{Vent-Prior's Letter of Introduction (Frozen Slate)} --- A slate tablet that must be held against warm skin to read. \texttt{[SOCIAL]} \texttt{[AUTHORITY]}
  \hfill \textit{Grants access to any Karak-Aeler mining hold. The slate is cold. It grows colder when the holder lies.}
\item \textbf{Demon's Nail (Iron Splinter)} --- A piece of the Ninth Seal's timber. \texttt{[MAGIC]} \texttt{[STEALTH]}
  \hfill \textit{Touch it to a lock; the lock opens. The lock also screams. The scream alerts the demon. The nail crumbles after one use. The demon knows your face now.}
\item \textbf{Ice-Lama's Blessing (Frozen Tear)} --- A drop of water that never melts, said to be the saint's tear. \texttt{[HEALING]}
  \hfill \textit{Drink it to remove one level of Harm caused by cold or exhaustion. The tear tastes of salt and memory.}
\item[J] \textbf{The Ninth Echo's Whisper (Slate)} --- A piece of slate carved with the ninth echo's words. \texttt{[COMBAT]} \texttt{[FEAR]}
  \hfill \textit{Play it to silence a foe. The foe hears their own name and freezes for one exchange. The slate then crumbles. The echo is now silent. The bridge has one fewer whisper.}
\item[Q] \textbf{First Keystone's Echo (Tuning Fork)} --- A tuning fork that rings with the mountain's voice. \texttt{[LEGEND]}
  \hfill \textit{Strike it, and the mountain will answer one question --- but the answer is always a riddle, and the riddles are always about obligations.}
\item[K] \textbf{High Folk Novice's Memory (Bone Needle)} --- A needle that the novice used to sew her silence. \texttt{[SECRET]}
  \hfill \textit{Thread it with a strand of your own hair and sew a wound closed. The wound heals, but you forget why you received it. The needle remembers.}
\item[A] \textbf{Vent-Prior's Forgiveness (Iron Token)} --- A token that grants pardon for any crime in the Himadri. \texttt{[AUTHORITY]}
  \hfill \textit{The pardon is conditional. The condition is a secret mission: re-seal the demon permanently. The Vent-Prior's bell will not ring for this lie.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:himadri-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 -> 4, 6--10 -> 6, J/Q/K -> 8, A -> 10).
\item Diamonds are codified outcomes (bells, ingots, prayers) that change position rather than call for a roll.
\item If any A appears, echo \textbf{mountain-omens}---iron that weeps rust, bells that ring without clappers, a pass that closes behind you, a whisper that sounds like your own breath, a shadow that walks without a body.
\end{itemize}

\subsection*{Additional Features (High Altitude Mechanics)}
\label{sec:himadri-mechanics}

\begin{itemize}
\item \textbf{Breath-Counting:} In high passes (above 4,000 metres), characters must mark Fatigue for every hour of exertion. A character who deliberately paces their breathing (making a Resolve test, DV 3) can reduce the Fatigue cost by 1. Counting aloud to nine in any context triggers a \emph{Mountain's Attention} roll (Spirit + Resolve, DV 4). On failure, the mountain takes something: 1 Fatigue, a minor memory (erase a fact), or a shadow (the character casts no reflection until they leave the high passes).
\item \textbf{The Ninth Seal Clock [8]:} The Covenant seal on the demon mine has a hidden timer that ticks when the demon whispers, when someone takes iron from the sealed tunnel, or when a breath-promise is broken nearby. When the timer fills, the seal cracks. The party can reset the timer by reinforcing the seal with a new Covenant rite (costs a memory) or by feeding the demon a story (costs a truth). The demon prefers stories about betrayal. It pays attention to those.
\item \textbf{Mountain Offerings:} The High Folk believe that the mountain must be fed with offerings of silence, memory, or warmth. A character who leaves an offering (a pinch of salt, a whispered secret, a lock of hair) at a shrine or a cairn may roll Spirit + Lore (DV 3). On success, the next weather hazard in that pass is downgraded (blizzard to snow, ice to frost). On failure, the offering is rejected --- the character gains the \emph{Marked by the Mountain} condition (the next environmental complication is automatically Desperate).
\item \textbf{High Folk Taboos:} The High Folk have many taboos: never point at a peak, never speak the name of a dead person in a pass, never refuse a cup of butter tea, never step over a prayer flag. Breaking a taboo in front of a High Folk witness causes the witness to fall silent and refuse to help. The GM may spend 1 SB to have a broken taboo attract the attention of a mountain spirit (a complication from the \emph{Clubs} table).
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Adur the Forge-Father} --- Oath of Flame and Light. \textit{The fire in the smelter, the heat that tempers iron and oaths.}
    \item \textbf{Hemrir the Wanderer} --- The Sealed Gate. \textit{The keeper of thresholds; his shrines are bridge-posts and tunnel mouths.}
    \item \textbf{The Breath-Eater} --- Khemesh. \textit{The pressure in the deep; the silence that follows a cave-in.}
    \item \textbf{The Ice-Lama} --- The Pale Shepherd. \textit{The saint who froze to death; she guides travellers through whiteouts.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Ninth Seal and the Kalrantha Amulet (Dhahara):} The demon in the sealed mine is a fragment of Varash the Unmaker, the same entity trapped in the Kalrantha amulet. The Veiled Sisters want to free this fragment to use against the Covenant. The Breath-less want to re-seal it permanently. The demon wants to be whole. It offers favours to anyone who brings it a shard of green glass.
  \item \textbf{Iron Trade and the Ngomebe Stone Men:} The iron from the Himadri mines is shipped to Ngomebe to forge rollers for the walking cities. A disruption in the iron supply could stop a city mid-migration. The Ngomebe Mason-Kin have a permanent envoy at the Vent-Prior's hall. The envoy is a woman who speaks only in stone-songs. She has never smiled.
  \item \textbf{Breath-Promises and the Covenant (Mykkiel):} The Covenant recognizes Aeler breath-promises as legally binding, but only if witnessed by a judge. The Seal-Keeper is the only Covenant authority in the high passes. He has not witnessed a new promise in twenty years. He says the mountain remembers all obligations anyway.
  \item \textbf{The Silent Pass and the Hollow (Eastern Realms):} The Silent Pass is a known threshold where the Hollow's attention is thin. Aelaerem hedge-keepers leave offerings of salt at the pass entrance to keep the Oath-Keepers from following caravans. The salt must be from the pans of the Silent Widows. Any other salt will attract them. The Hedge-Keepers will not explain why. They say the salt knows.
\end{itemize}

\subsection*{Himadri Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Mine:} Crumble a crust into the tunnel mouth. Ignore the first \emph{Whiteout} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Bell Rope:} Tie a knot in the rope of the shrine bell. Cancel \emph{Ninth Echo} or take --1 die on your next stealth roll (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Iron:} Whisper a dead miner's name into an ingot. Gain +1 Effect when negotiating with Karak-Aeler this scene, but mark \emph{Promise} to the Vent-Prior. \texttt{[SUPERSTITION]}
  \item \textbf{Butter to the Shrine:} Leave a lump of yak butter at a cairn. The next time you cross a pass, the mountain will not take your shadow. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:himadri-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Karak-Aeler mine-captains or High Folk lamas}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{watchers or yak-herders}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{mountain spirits or demon-touched miners}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favourable weather}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local mine-captain or seal-keeper} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{whiteout or crevasse field}.
\item \textbf{Foe Appears:} A \textbf{demon-touched miner or a territorial snow cat} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Karak-Aeler miners, High Folk watchers, or demon fragments} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes --- \textbf{seal cracks / ninth echo spoken / whiteout worsens}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Vent-Prior, Lopön, or Demon from the Ninth Seal} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically --- \textbf{mine collapses / bridge falls / ice shrine cracks}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice --- \textbf{omen appears / blessing lost / curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Himadri (High Altitude):} Breath-counters tick backward, snow falls upward, the air tastes of copper.
\item \textbf{Himadri (Demon Seal):} The Ninth Seal's glow intensifies, demon whispers become audible, a crack appears in the clay.
\item \textbf{Himadri (High Folk):} A prayer flag tears, a yak lows at midnight, a lama's bell rings without being struck.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Frostbite}.
\item \textbf{Cut the Price (Presence + Sway):} On a hit, freeze range for one exchange; if a price waiver is in your possession, +1 die.
\item \textbf{Weather the Whiteout (Resolve + Athletics):} On a hit in \emph{Whiteout}, you pick who finds shelter; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Breath \& Demon)}
\begin{itemize}
\item \textbf{The Breath-Eater (Nine Silences)} --- \emph{Phases}: Mines Collapsing $\rightarrow$ Orchestrated Breath-Promise Break $\rightarrow$ Pass Coup. \emph{Cracks}: find a lost breath-counter, expose a false keystone, survive a count of nine.
\item \textbf{The Unsealed (Demon of the Ninth Seal)} --- \emph{Phases}: Whispers in the Mine $\rightarrow$ Procession of Cracking Seals $\rightarrow$ Iron-Baptism of a Vent-Prior. \emph{Cracks}: relic Covenant seal, survivor testimony, the mine's face weeping clear water.
\item \textbf{The Ice-Lama's Waking (High Folk Saint)} --- \emph{Phases}: Frozen Tears $\rightarrow$ The Cold Remembers $\rightarrow$ The Ninth Breath Spoken. \emph{Cracks}: return a stolen prayer flag, speak a true regret at her feet, survive a night in her cave without fire.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In the Himadri foothills, power flows through breath, silence, and the weight of the mountain's attention. The Karak-Aeler hold the mines and the forges, but they do not rule the High Folk --- they coexist with them, bound by ancient compacts sealed in frozen iron. The High Folk's lamas speak for the mountain spirits, and the spirits are not kind. The Covenant's seal on the Ninth Seal Mine is a reminder of what happens when the mountain is not respected. The demon whispers. The mountain listens. And the passes remember every breath taken in them.

\textbf{For the GM:}  
Power in the Himadri is measured in altitude, in the number of breaths you can hold, and in the stories the mountain chooses to remember. Patronage often takes the form of breath-counters, iron ingots, or prayer flags --- each tied to a specific pass, a specific spirit, a specific obligation. To emphasise the harsh, remote character:
\begin{itemize}
\item Make \textbf{breath-counting} a constant tension: characters must manage their Fatigue carefully. The mountain punishes haste.
\item Let the \textbf{Ninth Seal Clock} be a visible campaign front. The party's actions tick the timer; when it fills, the demon manifests. The demon does not fight --- it negotiates. It offers warmth, iron, safe passage. It asks for small betrayals. The High Folk say that the demon's bargains always leave a frost on the heart.
\item Use the \textbf{High Folk taboos} to create social tension. Breaking a taboo is not a crime --- it is an insult to the mountain. The High Folk will not explain the taboos beforehand; they expect visitors to watch and learn. A character who makes an offering without being asked gains a small boon. A character who refuses an offering gains the mountain's attention.
\end{itemize}
In the Himadri, the mountain forgets nothing --- and the ninth breath is always waiting.
\end{tcolorbox}

% Index entries
\index{Himadri Foothills|see{Roof of the World, Karak-Aeler}}
\index{Theme generator}
\index{Karak-Aeler generator}
\index{Places!Himadri}
\index{People!Himadri}
\index{Complications!Himadri}
\index{Rewards!Himadri}
\index{Karak-Aeler}
\index{High Folk}
\index{Ninth Seal demon}
\index{Breath-counting}
\index{Ice-Lama}
\index{Vent-Prior Khardung}
\index{Lopön Tenzin}
\index{Yak-Mother Pema}
\index{Local Patrons!Himadri}