\section{The Himadri Foothills --- ``The Iron Passes''}
\label{chap:himadri}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The mountain negotiates. The seal remembers. The ninth breath is never yours.}

\begin{quote}
  \textbf{Vent-Prior of the Iron Road (Elite):}  
  ``The mountain does not speak in words. It speaks in the crack of a bearing, the echo of a hammer, the taste of ore on the tongue. We Aeler do not conquer stone. We negotiate with it. The lease is not a document. It is the weight of the keystone that holds the roof.''
\end{quote}

\begin{quote}
  \textbf{Mine-Breath (Commoner):}  
  ``I carry a bell that counts my breaths. Eight is safe. Nine is the mountain's. My grandfather counted to nine once. The mountain took his shadow. He lived another thirty years, but he never cast a shadow again.''
\end{quote}

\textbf{Genre / Mood:} Industrial horror, mountain diplomacy, demonic containment.

\textbf{Starting Location:} The entrance to the Ninth Seal Mine, where a Covenant seal of clay and iron glows faintly in the dark. A vent-prior in a brass breath-counter steps forward: ``The seal is failing. The demon whispers through the crack. The rajas want to open it — they think the demon's iron can win their wars. The Covenant wants to re-seal it. The Aeler want to mine around it. You are here because you do not know enough to choose a side. That makes you useful.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Himadri foothills are the rugged transition between the Sarpies plateau and the towering peaks of the Aelerian mountains. Here, the air is thin and cold, the rivers run white with snowmelt, and the passes are measured in breath‑counts rather than miles. Iron mines scar the hillsides, worked by Dhaharan labourers and overseen by Aeler vent‑priors. Elephant caravans haul ore down to the Brass Coast; mule trains carry silver and tea up to the mountain holds. The Aeler have established a viceroyalty here — not conquest, but a long lease backed by breath‑bonds and keystone wax. The local rajas pay tribute in iron and receive protection from lowland enemies. It is an uneasy peace, written in bell‑codes and toll‑ledgers. The Aeler engineers speak a creole of their own tongue; the Dhaharan nobles wear Aeler breath‑counters as jewellery; and deep beneath the oldest mine, a demon sleeps behind a Covenant seal — a fragment of the thing that turned the Red Desert to glass.

\vspace{2mm}
\textbf{What you notice first:} The ringing of iron hammers and the lowing of elephants. The way every bridge has a dwarf toll‑keeper with a brass scale. The silence after a bell rings — the silence is for the mountain to answer.

\textbf{The one rule that kills newcomers:} Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one — a breath, a memory, a shadow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Keystone}

Power here is measured in iron quotas, breath‑bonds, and the age of a keystone. The Aeler do not rule through armies; they rule through the maintenance contracts that keep the mines from collapsing, the bridges from falling, and the passes from freezing. A vent‑prior's authority rests on his ability to read the mountain's cracks; a mine‑captain's on her skill with a tuning fork. The Dhaharan rajas pay tribute in ore, but they also hold the local rituals that appease the mountain spirits. The Covenant's sealed mine — the Ninth Seal — is a reminder of what happens when the mountain is not respected. The Aeler have kept it closed for three centuries. The seal is cracking.

\subsection*{Prominent NPCs of the Himadri Foothills}
\label{sec:himadri-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Vent-Prior Thrain & Aeler governor of the Iron Passes & Maximise ore output while keeping the demon sealed & He wears a bell around his neck; it rings when he lies. \\
\hline
Mine-Captain Korra & Aeler engineer & Keep the tunnels safe and the quotas met & She counts her breaths in eights; she has never reached nine. \\
\hline
Rani Jaya & Dhaharan raja of the eastern hills & Balance tribute to Aeler with loyalty to her clan & She lost an arm to a mine collapse; the prosthetic is a pickaxe. \\
\hline
Seal-Keeper Haran & Covenant priest of the Ninth Seal & Ensure the demon never wakes & He has not left the mine entrance in forty years. His eyes are rust-coloured. \\
\hline
Elephant Master Tarkan & Caravan leader & Move iron to the coast before the snows close the passes & He speaks to his elephants in Dhaharan; they answer in Aeler. \\
\hline
\end{longtable}

\subsection*{Names of the Himadri Foothills}
\label{sec:himadri-names}

A mix of Aeler gutturals (short, hard consonants) and Dhaharan vowels. Surnames may be Aeler clan names or Dhaharan hill titles.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/House} & \textbf{Epithets} \\
\hline
Thrain & Korra & Ironfoot & \textit{the Bell-Breaker} \\
\hline
Borin & Jaya & Stonehand & \textit{Breath-Counter} \\
\hline
Harkan & Sita & Aezelbron & \textit{Bridge-Warden} \\
\hline
Rurik & Laxmi & Thrainsdottir & \textit{Toll-Eater} \\
\hline
Vikram & Meera & Keystone-bearer & \textit{the Unshadowed} \\
\hline
\end{longtable}

\paragraph*{(Mine/Bridge/Seal)} Iron mine with a carved face above the entrance; Aeler toll bridge with a dwarf and a brass scale; the Ninth Seal Mine entrance with a Covenant seal of clay and iron.

\section*{Spades --- Places}
\label{sec:himadri-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Iron Mine of the Crying Face} --- A deep shaft with a carved face above the entrance.
  \hfill \textit{The face weeps rust. The tears are iron oxide. The mountain is mourning its dead.} `[INDUSTRIAL]`
\item \textbf{Aeler Toll Bridge} --- A stone arch over a gorge; a dwarf with a scale sits at the far end.
  \hfill \textit{The toll is not in coin — it is in weight. Your truth is weighed against a bar of iron.} `[AUTHORITY]`
\item \textbf{Ninth Seal Mine (Sealed)} --- A tunnel mouth sealed with Covenant clay and iron bands.
  \hfill \textit{The seal is warm. It glows faintly at night. The demon whispers through the cracks.} `[DEATH]` `[MAGIC]`
\item \textbf{Keystone Chamber} --- A cave where the Aeler store their breath-bond ledgers.
  \hfill \textit{The ledgers are carved on slate. Each slate is a keystone. Break one, void a contract.} `[AUTHORITY]`
\item \textbf{Silent Pass} --- A high mountain pass where no bird flies.
  \hfill \textit{The air is too thin for sound. Arguments must be written in the snow. The snow does not lie.} `[WEATHER]`
\item \textbf{Elephant Corral (Highland)} --- A fenced area with chains; elephants chained by one leg.
  \hfill \textit{The chains have names — each is the elephant's first kill. The names are in Aeler. The elephants remember.} `[LABOR]`
\item \textbf{Smelter's Forge (Aeler Mobile)} --- A wagon-mounted furnace that follows the ore caravans.
  \hfill \textit{The forge is pulled by a team of oxen. It is always last in the procession. The oxen have no shadows.} `[INDUSTRIAL]`
\item \textbf{Bridge of Nine Echoes} --- A rope bridge over a chasm; each step echoes nine times.
  \hfill \textit{The ninth echo is always a whisper. It is the name of the next person to fall. Do not answer.} `[DEATH]` `[FEAR]`
\item \textbf{Vent-Prior's Archive} --- A subterranean vault of breath-bonds and toll receipts.
  \hfill \textit{The archive is sealed with Aeler keystone wax. The wax is mixed with miner's blood. It never melts.} `[AUTHORITY]`
\item[J] \textbf{Abandoned Shaft (Demon's First Awakening)} --- A collapsed tunnel where the demon first spoke.
  \hfill \textit{The air is warm. The walls are coated in green glass — like the Kalrantha amulet. The glass sings at midnight.} `[MAGIC]` `[DEATH]`
\item[Q] \textbf{Shrine of the Ninth Breath} --- A small stone altar where miners leave offerings.
  \hfill \textit{The offerings are not coin — they are memories written on strips of leather. The leather never rots. The memories are still warm.} `[RITUAL]`
\item[K] \textbf{Raja's Hill Fort} --- A stone fortress overlooking the main mine.
  \hfill \textit{The fort has a fire shrine and an Aeler breath-counter on the gate. Both are watched. The watchers do not blink.} `[POLITICS]`
\item[A] \textbf{The First Keystone (Myth)} --- A stone that never cracks, said to lie at the deepest point of the oldest mine.
  \hfill \textit{It is the mountain's heart. Touch it, and the mountain will ask a question. Answer poorly, and the passes close forever. Answer well, and it may ask another.} `[LEGEND]`
\end{enumerate}

\paragraph*{(Captain/Judge/Master)} Mine-Captain Korra with a tuning fork; Rani Jaya with a pickaxe prosthetic; Elephant Master Tarkan with a bone whistle.

\section*{Hearts --- People \& Factions}
\label{sec:himadri-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Mine-Captain (Aeler)} --- A dwarf with a tuning fork and a level. She can hear a crack in a beam from a hundred paces.
  \hfill \textit{She never walks under a keystone. The mountain has ears. She has never lost a crew.} `[AUTHORITY]`
\item \textbf{Rani Jaya} --- A young woman in armour, riding a caparisoned elephant. She lost an arm to a mine collapse.
  \hfill \textit{The prosthetic is a pickaxe. She can write with it. The letters are iron filings. Her clan trusts her. The Aeler fear her.} `[POLITICS]` `[COMMUNITY]`
\item \textbf{Seal-Keeper Haran (Covenant)} --- An old man in grey robes, sitting by the Ninth Seal. His eyes are rust-coloured.
  \hfill \textit{He has not eaten solid food in forty years. The seal's glow sustains him. He has forgotten his own name. The seal remembers.} `[JUSTICE]`
\item \textbf{Elephant Master Tarkan} --- Old, missing two fingers, tattooed with elephant footprints.
  \hfill \textit{He knows the name of every elephant that has crossed the pass in fifty years. The elephants know his. They call him ``Brother of the Long Road.''} `[COMMUNITY]`
\item \textbf{Vent-Prior Thrain} --- A dwarf with a bell around his neck. He stands behind the iron throne.
  \hfill \textit{The bell rings when he lies. It rings often. He is always lying about the ore quotas. The miners know. They work anyway.} `[AUTHORITY]`
\item \textbf{Aeler Toll-Keeper} --- A dwarf with a brass scale and a ledger. He weighs your goods and your lies.
  \hfill \textit{The scale tips for both. A lie makes the pan dip lower. The ledger is written in breath. It fogs when you are honest.} `[AUTHORITY]`
\item \textbf{Smelter Worker (Dhaharan)} --- A young man with red-rimmed eyes. He coughs iron dust.
  \hfill \textit{His spittle stains the ground. He will not live past thirty. He knows. He sends half his wage home to a village that has already forgotten him.} `[LABOR]` `[COMMUNITY]`
\item \textbf{Treasure Hunter (Foreign)} --- A foreigner digging in an old landslide. He is looking for the tomb of a prince.
  \hfill \textit{The prince is the one who stole the Kalrantha amulet. The hunter has a map tattooed on his back. The map moves.} `[GREED]`
\item \textbf{Breath-Bound Miner} --- A woman who sold her breath to the mountain for a higher quota. She can only speak in whispers.
  \hfill \textit{Her voice is a rustle. The mountain speaks through her. The other miners leave offerings of silence.} `[MAGIC]` `[DEBT]`
\item[J] \textbf{Ghost of the Landslide} --- A figure in grey, seen only at dusk. It points at the pass.
  \hfill \textit{It is pointing at a safe route — or a trap. No one has survived to find out. The ghost does not speak. It does not need to.} `[DEATH]`
\item[Q] \textbf{Vent-Prior's Cipher Clerk} --- A dwarf who writes diplomatic letters in code.
  \hfill \textit{The code is a poem. The poem is about a lost mine. The mine is the Ninth Seal. The clerk dreams in couplets.} `[AUTHORITY]`
\item[K] \textbf{Demon's Whisper (from the Ninth Seal)} --- A voice heard in the sealed mine.
  \hfill \textit{It offers a deal: a year of good fortune for a single memory. The memory is the last breath of a dying friend. The voice sounds like someone you loved.} `[MAGIC]` `[FEAR]`
\item[A] \textbf{The First Miner (Legend)} --- A skeleton holding a tuning fork, sitting at the First Keystone.
  \hfill \textit{If you take the tuning fork, you become the mountain's new voice. You will never leave the depths. The miners will hear you sing. They will not thank you.} `[LEGEND]`
\end{enumerate}

\paragraph*{(Cave-in/Toll/Echo)} A cave-in traps miners — the collapse was sabotage; the Aeler toll-keeper demands double payment; the ninth echo on the Bridge of Nine Echoes speaks a name.

\section*{Clubs --- Complications/Threats}
\label{sec:himadri-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Cave-In (Sabotage)} --- A tunnel collapses. The collapse was not an accident — a saboteur loosened the timbers. `[DISASTER]`
  \hfill \textit{The saboteur is in the party. Or the party is accused. The Mine-Captain is watching. Her tuning fork is warm.}
\item \textbf{Toll Dispute} --- The Aeler keeper demands double; the raja's men refuse. The dispute is a trap. `[AUTHORITY]` `[POLITICS]`
  \hfill \textit{Both sides want the party to choose a side. Neither is honest. The bridge is the only crossing. The scale is balanced on a lie.}
\item \textbf{Ninth Echo} --- On the Bridge of Nine Echoes, someone counts the echoes. The ninth echo says a name. `[DEATH]` `[FEAR]`
  \hfill \textit{The name is the speaker's. They faint. The bridge remembers. The name is now a toll. The bridge will collect.}
\item \textbf{Landslide (Pass Blocked)} --- The pass is blocked. The alternative route is three days longer. `[TRAVEL]`
  \hfill \textit{The landslide was deliberate — a rival caravan master wants to delay you. His camels are ahead. His guilt is written in the mud.}
\item \textbf{Breath Shortage (High Altitude)} --- At high altitude, the party struggles. Mark Fatigue 1 per scene. `[WEATHER]` `[SURVIVAL]`
  \hfill \textit{Acclimatisation requires a rest day. The rest day will cost a toll. The toll is a breath-bond. The mountain is watching.}
\item \textbf{Spy in Camp} --- A worker is stealing iron samples. The spy is working for an Ashaani agent. `[POLITICS]` `[STEALTH]`
  \hfill \textit{The samples are for a new alloy. The alloy is a weapon. The weapon is meant to break the Ninth Seal. The spy has a family. They do not know.}
\item \textbf{Covenant Audit} --- Seal-Keeper Haran demands to inspect the party's cargo. `[JUSTICE]` `[AUTHORITY]`
  \hfill \textit{The cargo includes a relic from the Red Desert — a piece of green glass. It is humming. Haran's eyes are weeping rust.}
\item \textbf{Aeler Bell Code Change} --- The toll-keeper rings a different sequence; the party is now ``unregistered.'' `[AUTHORITY]` `[TRAP]`
  \hfill \textit{The change was ordered by the Vent-Prior. The Vent-Prior is testing your loyalty. Fail, and you are banned from the passes. The ban is carved in slate.}
\item \textbf{Elephant Sickness (Curse)} --- A disease spreads through the caravan. The disease is a curse from a local spirit. `[DEATH]` `[MAGIC]`
  \hfill \textit{The spirit was disturbed by the mine's expansion. It wants a sacrifice. The sacrifice is a memory. The elephants are coughing up names.}
\item[J] \textbf{Vent-Prior's Summons} --- Thrain demands the party attend him. He offers a deal: find the saboteur in exchange for a toll waiver. `[AUTHORITY]` `[POLITICS]`
  \hfill \textit{The saboteur is his own clerk. The clerk is his daughter. She is trying to open the Ninth Seal. She wants to free the demon. She thinks it will speak to her.}
\item[Q] \textbf{Silent Pass's Cold (Supernatural)} --- The temperature drops; the party loses 1 Supply per hour. `[WEATHER]` `[MAGIC]`
  \hfill \textit{The cold is supernatural — a frost spirit has taken residence. It wants a breath-bond. It will accept a story. The story must be about fire.}
\item[K] \textbf{Demon Wakes (Ninth Seal Crack)} --- The sealed mine cracks open. A shadow escapes; it possesses the nearest Aeler engineer. `[DEATH]` `[MAGIC]`
  \hfill \textit{The engineer begins to sabotage the pass. The demon wants the lease broken. The demon wants the mountain to weep. The engineer's face is not his own.}
\item[A] \textbf{First Keystone Quakes} --- The mountain's heart trembles. The passes close. The mines collapse. The demon is free. `[APOCALYPSE]`
  \hfill \textit{The party has one day to re-seal the demon or flee. The Vent-Prior is weeping. The seal-keeper is dead. The mountain is counting.}
\end{enumerate}

\paragraph*{(Bell/Ingot/Seal)} Aeler breath-counter (silver bell, wear to count breaths safely); iron ingot (stamped with the Vent-Prior's mark); Covenant seal (clay, for witnessing oaths in the passes).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:himadri-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Aeler Breath-Counter (Silver Bell)} --- Wear it to count your breaths in high altitude. The bell rings when you reach eight. `[TRAVEL]` `[SURVIVAL]`
  \hfill \textit{Stop at eight. The ninth breath is the mountain's. The bell also rings when a demon is near. The tone is flat.}
\item \textbf{Iron Ingot (Vent-Prior's Mark)} --- A single bar of high-grade iron, stamped with a dragon. Worth 1 Measure. `[TRADE]`
  \hfill \textit{The dragon is the Vent-Prior's mark. Present it to any Aeler forge for a favour. The favour is a breath-bond. The ingot is still warm from the mountain's heart.}
\item \textbf{Covenant Seal (Clay, Himadri)} --- A lump of clay. Heat it to witness any oath in the foothills. `[JUSTICE]` `[SOCIAL]`
  \hfill \textit{The heat comes from a coal taken from the Ninth Seal mine. The coal whispers. The seal lasts one use. The clay is porous. It drinks lies.}
\item \textbf{Toll Waiver (Brass Nail)} --- A brass nail. Present it to pass any Aeler bridge for free. `[TRAVEL]`
  \hfill \textit{The nail fits into a hole in the toll-keeper's scale. The hole is for bribes. The nail is also a key. It opens a lock in the Vent-Prior's archive.}
\item \textbf{Elephant Whistle (Bone)} --- A bone whistle. Blow it to call an elephant from the caravan. `[TRAVEL]` `[COMMUNITY]`
  \hfill \textit{The whistle is made from the tusk of a bull that died in a rockslide. Its ghost runs with the elephant. The elephant will come. It may bring friends.}
\item \textbf{Raja's Signet Ring (Silver)} --- A silver ring with a tiger. Commands loyalty from the raja's clan. `[SOCIAL]` `[POLITICS]`
  \hfill \textit{The ring is a copy. The real one was lost in a landslide. The ghost of the original owner wears it. He will ask for it back. He will be polite.}
\item \textbf{Smelter's Gloves (Asbestos)} --- Handle hot metal safely. The gloves are cursed: they itch when you lie. `[CRAFT]` `[SOCIAL]`
  \hfill \textit{The itch is mild at first. It becomes unbearable with repeated lies. The gloves also resist demonic whispers. The demon does not like asbestos.}
\item \textbf{Map of the Silent Pass (Leather)} --- A chart drawn on leather. Shows a safe route through the high pass. `[TRAVEL]`
  \hfill \textit{The map is incomplete. The missing section is a fold. The fold is a secret — a cave with a spring that never freezes. The spring tastes of iron. It is the mountain's blood.}
\item \textbf{Vent-Prior's Letter of Introduction} --- A sealed scroll. Grants access to any Aeler hold in the mountains. `[SOCIAL]` `[AUTHORITY]`
  \hfill \textit{The letter is written in blood. The blood is the Vent-Prior's. He changes his blood type each season. The clerks have a chart. They update it daily.}
\item[J] \textbf{Seal-Keeper's Prayer (Clay Tablet)} --- A tablet inscribed with the original Covenant sealing rite. `[MAGIC]` `[JUSTICE]`
  \hfill \textit{Recite it to strengthen the Ninth Seal for one year. The recitation costs a memory. The memory is forgotten. The seal will remember.}
\item[Q] \textbf{Demon's Nail (Timber from Ninth Seal)} --- A piece of the sealed mine's timber. Touch it to a lock; the lock opens. `[MAGIC]` `[STEALTH]`
  \hfill \textit{The lock also screams. The scream alerts the demon. The nail crumbles after one use. The demon knows your face now. It will whisper to you in dreams.}
\item[K] \textbf{Ninth Echo's Whisper (Slate)} --- A piece of slate carved with the ninth echo's words. Play it to silence a foe. `[COMBAT]` `[FEAR]`
  \hfill \textit{The foe hears their own name and freezes for one exchange. The slate then crumbles. The echo is now silent. The bridge has one fewer whisper.}
\item[A] \textbf{Vent-Prior's Forgiveness (Pardon)} --- A written pardon for any crime in the Himadri foothills. `[AUTHORITY]` `[SOCIAL]`
  \hfill \textit{The pardon is conditional. The condition is a secret mission. The mission is to re-seal the demon — permanently. The Vent-Prior's bell will not ring for this lie.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:himadri-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2–5 → 4, 6–10 → 6, J/Q/K → 8, A → 10).
\item Diamonds are codified outcomes (bells, ingots, nails) that change position rather than call for a roll.
\item If any A appears, echo \textbf{mountain-omens}—iron that weeps rust, bells that ring without clappers, a pass that closes behind you, a whisper that sounds like your own breath.
\end{itemize}

\subsection*{Additional Features (Aeler Mechanics)}
\label{sec:himadri-mechanics}

\begin{itemize}
\item \textbf{Breath-Bonds:} Aeler contracts are sealed with breath — the parties exhale into a brass counter, and the bond is recorded as a weight of air. Breaking a breath-bond causes the breaker to lose the ability to draw a full breath for a year (mark Fatigue 1 at the start of each scene). Breath-bonds are measured in eighths; the ninth part is always held by the mountain. A bond that is kept for nine years is said to become a \textit{mountain-friend} — the signatory can walk any pass without toll.
\item \textbf{Keystone Law:} Every Aeler mine has a keystone — a single block that bears the weight of the roof. The keystone is also a legal document. Carve a contract into it, and the contract is enforceable by the mountain itself. Defacing a keystone is punishable by burial in the mine. The keystone does not forget. It does not forgive. It may, however, negotiate.
\item \textbf{The Ninth Seal Clock [8]:} The Covenant seal on the demon mine has a hidden clock that ticks when the demon whispers, when someone takes iron from the sealed tunnel, or when a breath-bond is broken nearby. When the clock fills, the seal cracks. The party can reset the clock by reinforcing the seal with a new Covenant rite (costs a memory) or by feeding the demon a story (costs a truth). The demon prefers stories about betrayal. It pays attention to those.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Adur (the Forge-Father)} — Oath of Flame and Light. *The fire in the smelter, the heat that tempers iron and oaths.*
    \item \textbf{Hemrir the Wanderer} — The Sealed Gate. *The keeper of thresholds; his shrines are bridge-posts and tunnel mouths.*
    \item \textbf{The Procession} — Clockwork Monad. *The optimisation of the mine; the hunger for efficiency that eats tradition.*
    \item \textbf{The Breath-Eater} — Khemesh. *The pressure in the deep; the silence that follows a cave-in.*
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Ninth Seal and the Kalrantha Amulet (Dhahara):} The demon in the sealed mine is a fragment of Varash the Unmaker, the same entity trapped in the Kalrantha amulet. The Veiled Sisters want to free this fragment to use against the Covenant. The Breath-less want to re-seal it permanently. The demon wants to be whole. It offers favours to anyone who brings it a shard of green glass.
  \item \textbf{Iron Trade and the Ngomebe Stone Men:} The iron from the Himadri mines is shipped to Ngomebe to forge rollers for the walking cities. A disruption in the iron supply could stop a city mid-migration. The Ngomebe Mason-Kin have a permanent envoy at the Vent-Prior's hall. The envoy is a woman who speaks only in stone-songs. She has never smiled.
  \item \textbf{Breath-Bonds and the Covenant (Mykkiel):} The Covenant recognises Aeler breath-bonds as legally binding, but only if witnessed by a judge. Sister Gaura (from the Satrapies) often travels to the Himadri foothills to witness the largest contracts. She is afraid of the mines. She goes anyway. She says the light is better underground.
  \item \textbf{The Silent Pass and the Hollow (Eastern Realms):} The Silent Pass is a known threshold where the Hollow's attention is thin. Aelaerem hedge-keepers leave offerings of salt at the pass entrance to keep the Oath-Keepers from following caravans. The salt must be from the pans of the Silent Widows. Any other salt will attract them. The Hedge-Keepers will not explain why. They say the salt knows.
\end{itemize}

\subsection*{Himadri Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Mine:} Crumble a crust into the tunnel mouth. Ignore the first \emph{Cave-In} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Bell Rope:} Tie a knot in the rope of the shrine bell. Cancel \emph{Ninth Echo} or take –1 die on your next stealth roll (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Iron:} Whisper a dead miner's name into an ingot. Gain +1 Effect when negotiating with Aeler this scene, but mark \emph{Obligation} to the Vent-Prior. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:himadri-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Aeler mine-captains or Covenant seal-keepers}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{toll-keepers or elephant masters}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{raja's spies or demon-touched miners}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favourable toll rates}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local mine-captain or seal-keeper} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{landslide or thin air}.
\item \textbf{Foe Appears:} A \textbf{demon-touched miner or a rival raja's guard} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Aeler engineers, raja's soldiers, or demon fragments} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes—\textbf{mine seals/keystone cracks/ninth echo spoken}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Vent-Prior, Seal-Keeper, or Demon from the Ninth Seal} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically—\textbf{mine collapses/bridge falls/first keystone quakes}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice—\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Himadri (Aeler Law):} Breath-counters tick backward, keystones weep water, a tuning fork rings without being struck.
\item \textbf{Himadri (Seal Law):} The Ninth Seal's glow intensifies, demon whispers become audible, a crack appears in the clay.
\item \textbf{Himadri (Pass Law):} Snow falls upward, the ninth echo repeats a different name each time, a bridge appears where there was only chasm.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Thin Air}.
\item \textbf{Cut the Toll (Presence + Sway):} On a hit, freeze range for one exchange; if a toll waiver is in your possession, +1 die.
\item \textbf{Weather the Cave-in (Body + Athletics):} On a hit in \emph{Cave-In}, you pick who escapes first; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Breath \& Demon)}
\begin{itemize}
\item \textbf{The Breath-Eater (Nine Silences)} — \emph{Phases}: Mines Collapsing $\rightarrow$ Orchestrated Breath-Bond Break $\rightarrow$ Pass Coup. \emph{Cracks}: find a lost breath-counter, expose a false keystone, survive a count of nine.
\item \textbf{The Unsealed (Demon of the Ninth Seal)} — \emph{Phases}: Whispers in the Mine $\rightarrow$ Procession of Cracking Seals $\rightarrow$ Iron-Baptism of a Vent-Prior. \emph{Cracks}: relic Covenant seal, survivor testimony, the mine's face weeping clear water.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In the Himadri foothills, power flows through breath-bonds, iron quotas, and the authority of the keystone. The Aeler vent-priors hold the contracts that keep the mines open; the Dhaharan rajas hold the land; the Covenant holds the seal on the demon. No single faction can rule alone, because the mountain itself is a stakeholder — and the mountain has a vote.

\textbf{For the GM:}  
Power in the Himadri foothills is measured in breath-counts, toll receipts, and the age of a keystone. Patronage often takes the form of breath-counters, iron ingots, or seal fragments — each tied to a specific mine, pass, or contract. To emphasize the Aeler focus:
\begin{itemize}
\item Make \textbf{breath-bonds} a central mechanic: players can enter into bonds for favours, but breaking them has harsh consequences. A bond kept for nine years becomes a \textit{mountain-friend} — the holder can walk any pass without toll.
\item Let the \textbf{Ninth Seal Clock} be a visible campaign front. The party's actions tick the clock; when it fills, the demon manifests. The demon does not fight — it negotiates. It offers iron. It offers favours. It asks for small betrayals.
\item Use the mines, bridges, and the sealed tunnel as arenas for social contests, where a breath-count is a threat and a keystone is a judge. The mountain is always listening. It may offer advice. The advice is usually good. The price is never stated.
\end{itemize}
In the Himadri foothills, the mountain forgets nothing — and the ninth breath is always waiting.
\end{tcolorbox}

% Index entries
\index{Himadri Foothills|see{Iron Passes, Aeler}}
\index{Theme generator}
\index{Aeler generator}
\index{Places!Himadri}
\index{People!Himadri}
\index{Complications!Himadri}
\index{Rewards!Himadri}
\index{Aeler breath-bonds}
\index{Ninth Seal demon}
\index{Keystone law}
\index{Iron mines}
\index{Silent Pass}
\index{Vent-Prior}
\index{Mine-Captain}
\index{Covenant seal}
\index{Local Patrons!Himadri}