\section{Lethai-thora --- ``The Silent Ledgers of Sihai''}
\label{chap:lethai-thora}

\textbf{Key Demographics: } Lethai-thoraa
\textbf{Capital City: } Linshu

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{We do not burn the dangerous scrolls. We bury them on shelves where no one thinks to look. Silence is not emptiness. It is a door that opens only inward.}

\begin{quote}
  \textbf{\textit{Keeper of the Seventh Gutter (Elite):}}  
  \textit{``We do not burn the dangerous scrolls. We bury them in plain sight, on shelves where no one thinks to look. The Emperor's Censor once stood in our archive and asked for a tax ledger. We gave him one. He did not notice the shelf behind him held the true history of his own grandfather's death. Silence is not emptiness. It is a door that opens only inward.''}
\end{quote}

\begin{quote}
  \textbf{\textit{Scroll-Binder (Commoner):}}  
  \textit{``My grandmother tied this scroll with a silk cord. The cord has nine knots. The ninth knot is a secret she never told me. When I die, I will tie a tenth. Someone will untie it in a century. They will learn why the emperor's concubine wept on the night of the lantern festival. Some secrets are too heavy for one lifetime.''}
\end{quote}

\textbf{Genre / Mood:} Bureaucratic melancholy, hidden history, silence as power.

\textbf{Starting Location:} A hidden pavilion behind a waterfall, where scrolls are stacked to the ceiling and the air is dry as bone. A Keeper of the Seventh Gutter looks up from a brass lamp: \textit{``The Emperor's Quiet Chancellor has sent a messenger. He wants a scroll that does not exist. He wants the truth about his mother's death. The truth is written in water. It will vanish if we speak it. But you — you are not bound by our silence. Will you carry a memory for us?''}

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Lethai-thora were once courtiers of a celestial fey realm that spanned the Valewood and the eastern hills. When the old fey courts fell — when the Alder King withdrew and the Hazel Queen sealed her gates — many Lethai-thora fled east. They crossed the mountains with nothing but armfuls of scrolls and the memory of a civilisation that had forgotten how to die.

In the fertile river valleys of what would become Sihai, they found scattered agrarian tribes fighting over water and silence. The Lethai-thora did not conquer. They taught. They taught the tribes to write contracts instead of drawing blood, to measure grain instead of stealing it, to record debts instead of nursing grudges. Over generations, the tribes became a kingdom, the kingdom became an empire, and the empire became Sihai. The Lethai-thora became its memory.

Now they dwell in hidden pavilions, preserving \textit{forgotten ledgers} that record what the Mandate of Heaven prefers to forget. Their knowledge is a burden: each secret they preserve slowly erases a piece of themselves. An Lethai-thora who guards a forbidden truth may find that she can no longer recall the color of her mother's dress — a small price, she says, for keeping the ledger whole. They are no longer fey. They are no longer fully human. They are something in between: scribes of a story that has no ending, archivists of a past that never quite happened.

\vspace{2mm}
\textbf{What you notice first:} The smell of old parchment and incense. The way scrolls are stacked with their spines facing the wall — so the words cannot see who reads them. The silence when a secret is spoken; even the insects stop chirping.

\textbf{The one rule that kills newcomers:} Never touch a scroll without announcing your lineage. Even a false lineage is better than silence. The scrolls remember who opened them, and they will whisper your name to the Hollow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Forgotten}

Power here is measured in scrolls sealed, secrets kept, and the length of a silence. The Lethai-thora have no king, no army, no treasury. They have archives, and the archives are their fortress. The Emperor's Censorate fears them because they cannot audit a secret that has no ink. The Mandate of Heaven is written in public; the Lethai-thora keep the unwritten ledger — the one that records why dynasties fall. They are not rebels; they are witnesses. And a witness who never speaks is the most dangerous kind.

\subsection*{The Fledging of the Empire (Origin Story)}
\label{sec:lethai-thora-origin}

When the Valewood's fey courts crumbled, a caravan of Lethai-thora refugees crossed the eastern passes. Their scrolls were water-stained, their children hungry, their memories already fraying. They found a patchwork of warring clans along the Sihon River — farmers who fought over silt, herders who stole each other's sheep, chieftains who could not agree on a name for the season.

The Lethai-thora did not offer swords. They offered ink.

They taught the clans to write their disputes on clay, to seal them with witnesses, to let the record speak instead of the spear. A generation later, the first emperor rose — a man who had learned his letters from a Lethai-thora tutor. He did not forget. He built his capital around the Lethai-thora archive, and he decreed that no law could be written without a copy placed in their care.

The Lethai-thora have been part of Sihai ever since — embedded so deeply that they can no longer imagine leaving, even if they wanted to. They are the empire's memory, but also its conscience, its secret‑keeper, and its silent judge. Some among them resent this entanglement; they dream of returning to the Valewood, of hearing the old songs again. But the Valewood has changed. The fey courts do not answer. And the Lethai-thora have become something they never intended: the architects of a mortal empire.

\subsection*{Prominent NPCs of the Lethai-thora}
\label{sec:lethai-thora-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Keeper Lian & Master of the Seventh Gutter & Preserve the true history of the emperor's grandfather & She has not spoken aloud in twenty years. She writes on water. \\
\hline
Scribe Wei & Scroll-binder & Tie a secret so tightly it can never be untied & He has forgotten his own name. He signs his work with a thumbprint. \\
\hline
Ink-Maker Hana & Forger of context keys & Create inks that reveal different truths in different seasons & She has a jar of ink made from crushed moonstones. It only writes during eclipses. \\
\hline
The Quiet Chancellor & Imperial censor (secret ally) & Find the scroll that proves the emperor's claim is false & He speaks in couplets; the second couplet is always a bribe. \\
\hline
The Ghost of the Ninth Scroll & A memory that was never written & A secret that killed its keeper before it could be recorded & It appears as a blank scroll that weeps ink. The ink is the secret. \\
\hline
\end{longtable}

\subsection*{Names of the Lethai-thora}
\label{sec:lethai-thora-names}

Inspired by the central kingdom traditions: soft consonants, vowel endings, honorifics. Surnames denote an archive, a gutter, or a lineage.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/House} & \textbf{Epithets} \\
\hline
Lian & Hana & Seventh Gutter & \textit{the Silent} \\
\hline
Wei & Mira & Forbidden Archive & \textit{the Unwitnessed} \\
\hline
Chen & Nuan & Ink-Thread & \textit{Ninth Scroll} \\
\hline
Jin & Yue & Water-Pavilion & \textit{the Forgetful} \\
\hline
Liam & Sora & Root-Memory & \textit{the Context-Keeper} \\
\hline
\end{longtable}

\paragraph*{(Archive/Chamber/Shrine)} A hidden pavilion with scrolls stacked to the ceiling; a brass lamp that never goes out; a shrine to the forgotten gods, covered in dust.

\section*{\spadecard{J} Places}
\label{sec:lethai-thora-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Hidden Pavilion (Waterfall)} --- An archive behind a curtain of falling water.
  \hfill \textit{The water muffles sound. Secrets spoken here cannot be overheard — but the water remembers.} \texttt{[SECRECY]}
\item \textbf{Chamber of the Ninth Scroll} --- A locked room with a single blank scroll on a pedestal.
  \hfill \textit{The scroll is not blank. It is written in disappearing ink. The reagent is moonlight.} \texttt{[MYSTERY]}
\item \textbf{Well of Forgotten Names} --- A stone well where the Lethai-thora throw scrolls that are too dangerous to keep.
  \hfill \textit{The well is bottomless. The scrolls burn as they fall. The smoke forms faces.} \texttt{[DEATH]}
\item \textbf{Ink-Maker's Studio} --- A cave with grinding stones and jars of pigment.
  \hfill \textit{The ink is made from crushed beetles, iron ore, and tears. The tears are the ink-maker's own.} \texttt{[CRAFT]}
\item \textbf{Silent Scriptorium} --- A hall where scribes copy scrolls without speaking.
  \hfill \textit{They write with their left hands. Their right hands are for swearing oaths.} \texttt{[LABOR]}
\item \textbf{Context Key Archive} --- A vault of stones, each carved with a season and a place.
  \hfill \textit{A scroll cannot be read without the correct context key. The keys are scattered across Sihai.} \texttt{[TRAVEL]}
\item \textbf{Emperor's Censorate (Monitoring Station)} --- A building near the archive, staffed by imperial spies.
  \hfill \textit{They pretend not to know about the hidden pavilion. The Lethai-thora pretend not to notice them.} \texttt{[AUTHORITY]}
\item \textbf{Bridge of Unspoken Names} --- A stone arch where the Lethai-thora whisper secrets to the river.
  \hfill \textit{The river carries the secrets to the sea. The sea does not tell.} \texttt{[RITUAL]}
\item \textbf{Shrine of the Silent Gods} --- A ruined temple where the old fey beings were once worshipped.
  \hfill \textit{The gods are not dead. They are waiting. The Lethai-thora leave offerings of ink.} \texttt{[LEGEND]}
\item[J] \textbf{The Forgotten Ledger's Vault} --- A subterranean room with a single iron door.
  \hfill \textit{The door has nine locks. The keys are held by nine different keepers. The ninth keeper is dead.} \texttt{[SECRECY]}
\item[Q] \textbf{Memory-Thread Chamber} --- A room where Lethai-thora weave their hair into cords.
  \hfill \textit{Each thread contains a memory. Cutting a thread releases the memory — and kills the keeper.} \texttt{[DEATH]}
\item[K] \textbf{The Emperor's Secret Archive (Hidden)} --- A duplicate archive built by a paranoid emperor.
  \hfill \textit{It contains the same scrolls as the Lethai-thora's, plus the emperor's own confessions.} \texttt{[AUTHORITY]}
\item[A] \textbf{The First Pavilion (Myth)} --- Where the first Lethai-thora hid the first forbidden scroll.
  \hfill \textit{It is said to be invisible. Only those who have forgotten nine names can see it.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Keeper/Scribe/Chancellor)} Keeper Lian with a brass lamp; Scribe Wei with ink-stained fingers; the Quiet Chancellor with a sealed letter.

\section*{\heartcard{Q} People \& Factions}
\label{sec:lethai-thora-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Keeper of the Seventh Gutter} --- Old, silent, writing on water with a reed brush.
  \hfill \textit{``The words vanish,''} she says. \textit{``That is how I know they are true.''} \texttt{[JUSTICE]}
\item \textbf{Scroll-Binder (Wei)} --- A man with no shadow. He lost it to a scroll that ate his reflection.
  \hfill \textit{He ties silk cords around each scroll. The knots are the secrets.} \texttt{[CRAFT]}
\item \textbf{Ink-Maker (Hana)} --- A woman with stained hands and a jar of moonstone ink.
  \hfill \textit{``This ink only writes during an eclipse,''} she says. \textit{``The next eclipse is in nine days.''} \texttt{[MAGIC]}
\item \textbf{Quiet Chancellor} --- An imperial censor in gray robes. He speaks in couplets.
  \hfill \textit{The first couplet is a greeting. The second is a threat. The third is a bribe.} \texttt{[POLITICS]}
\item \textbf{Memory-Weaver} --- A woman with a loom made of human hair. She weaves memories into cloth.
  \hfill \textit{The cloth is soft. It whispers the memory when you touch it.} \texttt{[ART]}
\item \textbf{Imperial Spy (Undercover)} --- A young man posing as a scribe. He has a hidden seal.
  \hfill \textit{He is looking for a scroll that names the emperor's true father. He will pay.} \texttt{[DECEPTION]}
\item \textbf{Ghost of the Ninth Scroll} --- A blank scroll that floats through the archive at midnight.
  \hfill \textit{It weeps ink. The ink forms words. The words are a secret that killed its keeper.} \texttt{[DEATH]}
\item \textbf{Forbidden Scholar} --- A monk who was excommunicated for reading a scroll he should not have.
  \hfill \textit{His eyes are gone. He reads by touch. The scrolls burn his fingers.} \texttt{[TRAGEDY]}
\item \textbf{Silent Guard} --- A warrior who took a vow of silence to protect the archive.
  \hfill \textit{He communicates by writing on his own skin. His skin is full of scars.} \texttt{[COMBAT]}
\item[J] \textbf{The First Keeper (Ghost)} --- A translucent figure in scholar's robes, seen at the well.
  \hfill \textit{She asks: ``What secret do you wish to forget?'' The answer becomes a scroll.} \texttt{[MYSTERY]}
\item[Q] \textbf{The Censorate's Interrogator} --- A man with a brass scale. He weighs truth and lies.
  \hfill \textit{His scale is balanced. He has never found a truth heavy enough.} \texttt{[AUTHORITY]}
\item[K] \textbf{The Emperor's Dreamer} --- A woman who dreams the emperor's nightmares. She knows what he fears.
  \hfill \textit{She is kept in a locked room. The Lethai-thora want to free her.} \texttt{[MAGIC]}
\item[A] \textbf{The Forgotten God's Echo} --- A voice from the ruined shrine. It speaks in a language no one understands.
  \hfill \textit{The Lethai-thora say it is the last fey being. It is asking for a memory.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Scroll/Forgery/Witness)} A scroll is found to be a forgery — the party is blamed; a context key is missing — the scroll cannot be read; the Quiet Chancellor demands a witness to a secret.

\section*{\clubcard{K} Complications/Threats}
\label{sec:lethai-thora-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Forged Scroll} --- A tax ledger is discovered to be a fake. The forgery is perfect. \texttt{[DECEPTION]}
  \hfill \textit{The forger is a Lethai-thora. The scroll is not a forgery — it is the original. The empire's copy is wrong.}
\item \textbf{Missing Context Key} --- A scroll cannot be read without a specific stone. The stone is lost. \texttt{[TRAVEL]}
  \hfill \textit{The stone was stolen by the Censorate. They want to destroy the scroll.}
\item \textbf{Witness Demanded} --- The Quiet Chancellor asks the party to witness a secret exchange. \texttt{[SOCIAL]}
  \hfill \textit{If they agree, they become bound by the Lethai-thora's silence. If they refuse, they are exiled from the archive.}
\item \textbf{Archive Fire (Accident)} --- A lamp falls. Scrolls burn. The smoke forms names. \texttt{[DISASTER]}
  \hfill \textit{The names are those of the keepers. One is yours.}
\item \textbf{Censorate Raid} --- Imperial inspectors arrive with a warrant. They have a list of forbidden scrolls. \texttt{[AUTHORITY]}
  \hfill \textit{The list is accurate. Someone inside the archive is a spy.}
\item \textbf{Memory-Thread Cut} --- A keeper's memory-thread is severed. She forgets everything. \texttt{[DEATH]}
  \hfill \textit{The thread was cut by an assassin. The assassin wants a secret only she knew.}
\item \textbf{Silence Broken} --- A keeper speaks aloud for the first time in decades. The Hollow hears. \texttt{[DEATH]}
  \hfill \textit{The Hollow sends an Oath-Keeper. The Oath-Keeper asks: ``What did you say?''}
\item \textbf{Ink of Truth (Forced Confession)} --- Someone poisons the ink well. Writing with it forces you to confess. \texttt{[MAGIC]}
  \hfill \textit{The poison was placed by the Censorate. They want to know where the forbidden scroll is.}
\item \textbf{Emperor's Death (Succession Crisis)} --- The emperor dies. The Mandate wavers. The archive is vulnerable. \texttt{[POLITICS]}
  \hfill \textit{The new emperor will burn the scrolls. The Lethai-thora must hide them — or die.}
\item[J] \textbf{Ghost of the Ninth Scroll Awakens} --- The blank scroll fills with ink. The ink is a name. \texttt{[DEATH]}
  \hfill \textit{The name is the next person to die. The name is yours.}
\item[Q] \textbf{The First Pavilion Appears} --- The mythic archive manifests. It is invisible to most. The party can see it. \texttt{[MYSTERY]}
  \hfill \textit{They are now witnesses to a secret that has been hidden for a thousand years. The Hollow will want it.}
\item[K] \textbf{Censorate's Interrogation} --- The party is arrested and brought before the brass scale. \texttt{[AUTHORITY]}
  \hfill \textit{The interrogator asks one question: ``What did the Lethai-thora tell you?'' The scale will know if you lie.}
\item[A] \textbf{The Forgotten God Wakes} --- The voice from the shrine becomes a scream. The ground shakes. \texttt{[LEGEND]}
  \hfill \textit{The god wants its memory back. The memory is held in a scroll. The scroll is the Ninth Scroll.}
\end{enumerate}

\paragraph*{(Scroll/Stone/Cord)} Forbidden scroll (one truth no one wants spoken); context key stone (unlocks one scroll's meaning); memory-thread (a cord containing a secret).

\section*{\diamondcard{A} Rewards/Leverage}
\label{sec:lethai-thora-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Forbidden Scroll} --- A scroll sealed with silk cord. It contains one truth no one wants spoken. \texttt{[TRUTH]}
  \hfill \textit{Read it aloud, and the secret becomes known. The speaker will be hunted.}
\item \textbf{Context Key Stone} --- A carved stone. Unlocks the meaning of one specific scroll. \texttt{[TRAVEL]}
  \hfill \textit{The stone is warm. It glows when held near the correct scroll.}
\item \textbf{Memory-Thread (Silk Cord)} --- A cord woven from a keeper's hair. Cut it to release a memory. \texttt{[MEMORY]}
  \hfill \textit{The memory is a secret. The keeper will forget it. The cutter will remember it.}
\item \textbf{Ink-Stone (Moonstone)} --- A jar of ink made from crushed moonstones. Writes only during eclipses. \texttt{[MAGIC]}
  \hfill \textit{The ink reveals truths that cannot be spoken. Use it to write a secret that will last.}
\item \textbf{Silence Token (Brass)} --- A small brass charm. Present it to demand a moment of silence in any court. \texttt{[SOCIAL]}
  \hfill \textit{During the silence, no one can lie. The truth will echo.}
\item \textbf{Keeper's Lamp} --- A brass lamp that never goes out. The flame flickers when a lie is spoken nearby. \texttt{[JUSTICE]}
  \hfill \textit{The lamp also flickers when a secret is hidden in the room.}
\item \textbf{Quiet Chancellor's Seal} --- A jade seal. Press it to any document to grant it imperial immunity from censorship. \texttt{[AUTHORITY]}
  \hfill \textit{The seal is a forgery. The Chancellor made it himself. It still works.}
\item \textbf{Scroll of Forgotten Names} --- A scroll that lists the names of everyone who has been erased from history. \texttt{[TRUTH]}
  \hfill \textit{Speak a name, and that person's deeds become known. The Hollow will remember them too.}
\item \textbf{Ink-Maker's Grinding Stone} --- A stone bowl. Grind any pigment into ink that writes the truth. \texttt{[CRAFT]}
  \hfill \textit{The truth cannot be erased. The ink will burn through lies.}
\item[J] \textbf{Ghost's Tear (Vial)} --- A vial of ink wept by the Ghost of the Ninth Scroll. \texttt{[DEATH]}
  \hfill \textit{Write a name with it, and that person will forget one secret. The secret is your choice.}
\item[Q] \textbf{First Pavilion's Key (Invisible)} --- An invisible key that opens the mythic archive. \texttt{[LEGEND]}
  \hfill \textit{You cannot see the key. You can feel it. It is cold.}
\item[K] \textbf{Emperor's Nightmare (Scroll)} --- A scroll containing the emperor's most feared secret. \texttt{[TRUTH]}
  \hfill \textit{Read it aloud to make the emperor confess — or to drive him mad.}
\item[A] \textbf{Forgotten God's Memory (Glass Shard)} --- A piece of green glass that hums. It contains a fragment of a fey god's memory. \texttt{[LEGEND]}
  \hfill \textit{Once, touch it to learn a secret of the old world. The secret will cost a memory of equal weight.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:lethai-thora-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2–5 \(\rightarrow\) 4, 6–10 \(\rightarrow\) 6, J/Q/K \(\rightarrow\) 8, A \(\rightarrow\) 10).
\item Diamonds are codified outcomes (scrolls, stones, threads) that change position rather than call for a roll.
\item If any A appears, echo \textbf{ink-omens}—scrolls that rustle in still air, a lamp that burns green, a forgotten name whispered from a shelf.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:lethai-thora-mechanics}

\begin{itemize}
\item \textbf{Silence as Sanctuary:} Once per scene, when asked a direct question, a character may spend 1 Boon to answer with silence. Treat Position as Dominant for resisting interrogation — but the GM gains 1 Story Beat (the silence has weight). In Lethai-thora halls, a silence is worth more than a thousand words.
\item \textbf{Context Keys:} To read an old text or recall a forbidden history, you need three keys: place, season, and witness. Without all three, roll at disadvantage. A Lethai-thora may create a temporary Context Key by spending 1 Boon and describing a small ritual (lighting incense, touching a stone, speaking a name).
\item \textbf{The Keeper's Toll:} When using a Lethai-thora Rite or consulting a hidden archive, mark 1 Fatigue. The knowledge costs something — a moment of warmth, a pleasant memory, the scent of rain. Some keepers have forgotten how to smile.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Unwritten} — The Witness. \textit{A secret that never needs to be spoken.}
    \item \textbf{The Quiet Scale} — Mykkiel. \textit{The law that weighs silence as heavily as testimony.}
    \item \textbf{The Ink Father} — Maelstraeus. \textit{The debt written in disappearing ink; the interest that accrues in forgetting.}
    \item \textbf{The Echo of the First Pavilion} — The Traveler. \textit{The path that leads only inward.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Ninth Scroll and the Kalrantha Amulet (Dhahara):} The Ninth Scroll is said to describe the creation of the amulet — and the demon sealed within. The Veiled Sisters have been searching for it for centuries. The Lethai-thora have hidden it in plain sight.
  \item \textbf{Silence Tokens and the Covenant (Mykkiel):} The Covenant recognizes a Lethai-thora silence token as a valid request for a closed hearing. A silence token can pause any court proceeding for the duration of a single breath.
  \item \textbf{Context Keys and the Ukwe (Sekogo):} The Ukwe's roots are a natural context key — they remember the season, the place, and the witness. A Lethai-thora who consults the Ukwe does not need to spend a Boon.
  \item \textbf{Memory-Threads and the Hollow (Eastern Realms):} The Hollow feeds on forgotten memories. A Lethai-thora's memory-thread is a feast. The Oath-Keepers will cut a thread if given the chance.
  \item \textbf{The Fledging of the Empire (Sihai):} The Lethai-thora's oldest scrolls still bear the water stains of the Valewood crossing. A keeper who unrolls one of these scrolls may glimpse a memory of the old fey courts — but the glimpse costs a year of her life.
\end{itemize}

\subsection*{Lethai-thora Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Archive:} Crumble a crust at the threshold. Ignore the first \textit{Archive Fire} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Scroll:} Tie a single knot in a silk cord. Cancel \textit{Forged Scroll} or take –1 die on your next forgery detection roll (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Well:} Whisper a dead keeper's name into the well. Gain +1 Effect when recalling a forgotten fact this scene, but mark \textit{Obligation} to the Lethai-thora. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:lethai-thora-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Censorate spies or imperial guards}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{scribes or keepers}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{memory-hunters or the Hollow}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or access to an archive}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local keeper or Censorate agent} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{flickering lamp or silent alarm}.
\item \textbf{Foe Appears:} A \textbf{Censorate interrogator or a Oath-Keeper} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Censorate agents, imperial guards, or Hollow spirits} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes—\textbf{scroll burns/thread cut/pavilion appears}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{High Keeper, Quiet Chancellor, or Ghost of the Ninth Scroll} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically—\textbf{archive floods/well overflows/first pavilion materialises}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice—\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Lethai-thora (Silence Law):} Lamps flicker without breeze, scrolls shuffle their own order, a whisper is heard from a sealed jar.
\item \textbf{Lethai-thora (Memory Law):} A keeper forgets your face, a memory-thread glows, you recall a fact you never learned.
\item \textbf{Lethai-thora (Context Law):} The season changes inexplicably, a stone context key warms, a witness appears who was not there.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth):} On a hit, shift one band. On a strong hit, also impose \textit{Silent Step}.
\item \textbf{Cut the Memory (Presence + Sway):} On a hit, freeze range for one exchange; if a memory-thread is in your possession, +1 die.
\item \textbf{Weather the Interrogation (Resolve + Spirit):} On a hit in \textit{Censorate's Interrogation}, you pick who is questioned; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Ink \& Silence)}
\begin{itemize}
\item \textbf{The Scroll-Eater (Nine Forgeries)} — \textit{Phases}: Scrolls Burned \(\rightarrow\) Orchestrated Erasure \(\rightarrow\) Archive Coup. \textit{Cracks}: recover a true scroll, expose a false context key, survive a night in the well.
\item \textbf{The Forgotten God (Memory Eater)} — \textit{Phases}: Voices from the Shrine \(\rightarrow\) Procession of Blank Scrolls \(\rightarrow\) Ink-Baptism of a Keeper. \textit{Cracks}: relic memory-thread, survivor testimony, the first pavilion opening its doors.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Lethai-thora, power flows through silence, ink, and the weight of a forgotten secret. A Keeper's authority rests on how many secrets she has buried and how many she has forgotten. The archive is the fortress; the silence is the shield. Yet they are also entangled with the empire they helped create — the Censorate watches them, the Emperor needs them, and the common people fear them. The Lethai-thora can no longer leave Sihai, even if they wished. Their fate is bound to the Mandate of Heaven.

\textbf{For the GM:}  
Power in Lethai-thora society is measured in scrolls sealed, context keys held, and the length of a pause. Patronage often takes the form of forbidden scrolls, context stones, or memory-threads — each tied to a specific secret, archive, or keeper. To emphasize the entangled relationship with Sihai:
\begin{itemize}
\item Use the \textbf{Silence as Sanctuary} mechanic to create tense social scenes where a pause is a weapon.
\item Require \textbf{Context Keys} to unlock ancient lore — the party must gather place, season, and witness before a scroll can be read.
\item Let the \textbf{Keeper's Toll} be a meaningful cost: fatigue, lost memories, or forgotten faces.
\item Remind players that the Lethai-thora are not outsiders; they are the empire's conscience. A Lethai-thora NPC may hold a secret about a PC's family, or about the true cause of a war, or about a crime the emperor himself committed.
\end{itemize}
In Lethai-thora society, the ink forgets nothing — and the silence remembers everything.
\end{tcolorbox}

% Index entries
\index{Lethai-thora|see{Silent Ledgers}}
\index{Theme generator}
\index{Silent Ledgers generator}
\index{Places!Lethai-thora}
\index{People!Lethai-thora}
\index{Complications!Lethai-thora}
\index{Rewards!Lethai-thora}
\index{Forbidden scrolls}
\index{Context keys}
\index{Memory-threads}
\index{Silence as sanctuary}
\index{Keeper's toll}
\index{Quiet Chancellor}
\index{Censorate}
\index{First Pavilion}
\index{Fledging of the Empire}
\index{Local Patrons!Lethai-thora}