\section{Lethai-ar (Ayokha)}
\label{chap:lethai-ar-ayokha-revised}

\textbf{Key Demographics: } Lethai-ar
\textbf{Capital City: } Nakhon Phanom

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The cord remembers. The river witnesses. The flame judges.}

\begin{quote}
  \textbf{Vigil Lady Suyin (Elite):}  
  ``A witness cord is not a leash. It is a mirror. If you speak truly, you see yourself. If you lie, you see what you could have been. I have cut two cords in my life. I still dream of the faces in the silk. The river remembers every word spoken on its banks. We are merely the scribes.''
\end{quote}

\begin{quote}
  \textbf{Wharf-Witness (Commoner):}  
  ``My lamp has burned for nine years without being refilled. The flame is green. It flickers when a merchant lies about the weight of his silk. The river does not need oil. It needs witnesses. I sit on this wharf and watch. That is my offering.''
\end{quote}

\textbf{Genre / Mood:} Riverine legal drama, oath-magic, communal memory, spirit diplomacy.

\textbf{Starting Location:} A monsoon wharf on the great river, where silk barges are being loaded. A Vigil in gray robes holds a lantern of green flame: ``The merchant and the captain have made a vow. The captain says the silk is eight bales. The merchant says it is nine. The witness cord is tied. It will tighten around the liar's wrist. I do not need to know who is lying. The cord will tell me. But you --- you are not bound. Watch. Learn. And do not pour the ninth cup.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Lethai-ar of Ayokha are the living bridges between mortals and the river spirits that carry the world's promises. Their ancestors were river-folk that ascended to the realm of the fair folk, and they now walk the middle path, a step between bodies of water and the world of men. They keep lanterns that never go out, record every contract spoken on the monsoon wharves, and bind their vows with silken cords dyed in river mud. Unlike their northern kin who swore to distant deities, the Lethai-ar serve \textbf{Benzaiten} (aligned with \textit{The Witness}) or \textbf{Inari} (aligned with \textit{The Traveler}). Their vows are physical cords, braided from their own hair and dyed with river mud. When a vow is fulfilled, the cord is burned. When it is broken, the cord tightens around the oath-breaker's wrist until blood is drawn (Harm 1). The Ayokhan taboo against the ninth cup is enforced by the Lethai-ar: the ninth cup is for the river spirits, and pouring it is an invitation to the Hollow.

\vspace{2mm}
\textbf{What you notice first:} The smell of river water and oiled silk. The way every Lethai-ar carries a small oil lamp, even at noon. The constant murmur of contracts being witnessed --- a merchant's promise, a captain's oath, a lover's vow.

\textbf{The one rule that kills newcomers:} Never pour the ninth cup of tea or sake in a Lethai-ar's presence. The ninth cup is for the witness. Pour it, and you have invited the river to judge you. The river is not merciful.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Witness Cord}

Power here is measured in lanterns kept lit, cords tied, and the number of oaths witnessed. The Lethai-ar have no army, no treasury, no territory. They have the river, and the river remembers everything. Every contract in Ayokha requires a Lethai-ar witness to be legally binding. The God-King's astrologers may set the calendar, but the Lethai-ar keep the chronicle. Their neutrality is absolute: they will witness a vow between enemies as readily as between allies. The price of a witness cord is not silver --- it is a promise to keep the river clean, to light a lamp at dusk, to speak no word that cannot be unsaid.

\subsection*{Prominent NPCs of the Lethai-ar (Ayokha)}
\label{sec:lethai-ar-npcs-revised}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Vigil Lady Suyin & Master of the central wharf & Keep the river's memory free of falsehoods & She has cut nine witness cords. The ninth was her own. \\
\hline
Cord-Maker Tama & Weaver of witness silk & Create a cord that cannot lie & She dyes her silk in river mud that glows under moonlight. \\
\hline
Lamp-Keeper Hiro & Guardian of the undying flame & Ensure the green lamp never goes out & He has not slept in nine years. The lamp keeps him awake. \\
\hline
River-Spirit (Nat) & Benzaiten's messenger & Collect stories from broken oaths & It appears as a woman with wet hair and riverweed. \\
\hline
The Oath-Broken (Exile) & A Lethai-ar who cut a false cord & Earn back her place in the Vigil & She carries a cord with nine knots. The ninth is cut. \\
\hline
\end{longtable}

\subsection*{Names of the Lethai-ar (Ayokha)}
\label{sec:lethai-ar-names-revised}

Inspired by the river kingdoms and monsoon coasts: soft, flowing consonants, honorifics (Lady, Master). Surnames denote a river, a wharf, or a cord.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/Epithet} & \textbf{Cord-Name} \\
\hline
Hiro & Suyin & River-watch & \textit{the Uncut} \\
\hline
Tama & Mira & Silk-hand & \textit{Ninth Knot} \\
\hline
Kaito & Yuki & Lamp-bearer & \textit{Green Flame} \\
\hline
Riku & Hana & Oath-strand & \textit{the Witness} \\
\hline
Sora & Nami & Mud-dyer & \textit{the Unforgotten} \\
\hline
\end{longtable}

\paragraph*{(Wharf/Temple/Cord-House)} A monsoon wharf with silk barges and green lanterns; a step-pyramid temple where the river spirit is honored; a cord-maker's house with hanging strands of dyed silk.

\section*{Spades --- Places}
\label{sec:lethai-ar-places-revised}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Monsoon Wharf (Central)} --- A stone pier where silk is loaded and oaths are witnessed.
  \hfill \textit{The pier is carved with the names of every oath kept. The names glow when a vow is fulfilled.} [\textbf{TRADE}] [\textbf{SOCIAL}]
\item \textbf{Step-Pyramid Temple (Benzaiten)} --- A stone pyramid rising from the riverbank. The top is a pool.
  \hfill \textit{The pool reflects not your face but the faces of those you have wronged.} [\textbf{SPIRIT}] [\textbf{TRUTH}]
\item \textbf{Cord-Maker's House} --- A wooden building with silk strands hanging from the rafters.
  \hfill \textit{The strands are dyed in nine shades of brown. The ninth shade is made from grave mud.} [\textbf{CRAFT}] [\textbf{MAGIC}]
\item \textbf{Lamp-Keeper's Shrine} --- A small stone hut with a single green flame.
  \hfill \textit{The flame has not gone out in three centuries. The keeper sleeps on a bed of ashes.} [\textbf{RITUAL}] [\textbf{COMMUNITY}]
\item \textbf{River Court (Floating)} --- A barge that serves as a mobile courthouse.
  \hfill \textit{The barge is anchored at a different spot each day. The river is the judge.} [\textbf{JUSTICE}] [\textbf{SOCIAL}]
\item \textbf{Vigil's Archive} --- A cave behind a waterfall, filled with witness cords.
  \hfill \textit{Each cord is labeled with a date and an oath. The oldest cord is nine centuries old.} [\textbf{MEMORY}] [\textbf{LEGEND}]
\item \textbf{Silk Village (Mulberry Groves)} --- A village where silk is spun and cords are woven.
  \hfill \textit{The mulberry trees were planted by the first Vigil. Their leaves are sacred.} [\textbf{COMMUNITY}] [\textbf{TRADE}]
\item \textbf{Bridge of the Ninth Clause} --- A wooden bridge where contracts are signed.
  \hfill \textit{The ninth clause is always the same: ``This oath is witnessed by the river.''} [\textbf{THRESHOLD}] [\textbf{LAW}]
\item \textbf{Well of Broken Oaths} --- A dry well where cut cords are thrown.
  \hfill \textit{The well is bottomless. The cords burn as they fall. The smoke smells of betrayal.} [\textbf{DEATH}] [\textbf{MEMORY}]
\item[J] \textbf{The Oath-Broken's Hut} --- A small shelter on the river's far bank, away from the wharves.
  \hfill \textit{The exile lives here. She tends a single lamp. The flame is blue, not green.} [\textbf{SOCIAL}] [\textbf{MYSTERY}]
\item[Q] \textbf{River-Spirit's Pool} --- A still pool behind the temple where Benzaiten's messenger appears.
  \hfill \textit{The pool is warm. It whispers. The whispers are the stories of broken oaths.} [\textbf{SPIRIT}] [\textbf{TRUTH}]
\item[K] \textbf{The God-King's Witness Chamber} --- A hall where the king's own oaths are recorded.
  \hfill \textit{The walls are covered in witness cords. The king has broken nine. The cords are still tight.} [\textbf{AUTHORITY}] [\textbf{JUSTICE}]
\item[A] \textbf{The First Cord (Myth)} --- A silk strand said to be woven from the river's first current.
  \hfill \textit{Whoever holds it can witness any oath anywhere. The price is a year of silence.} [\textbf{LEGEND}] [\textbf{TRUTH}]
\end{enumerate}

\paragraph*{(Vigil/Cord-Maker/Spirit)} Vigil Lady Suyin with a green lamp; Cord-Maker Tama with silk-dyed fingers; the River-Spirit with wet hair and riverweed.

\section*{Hearts --- People \& Factions}
\label{sec:lethai-ar-people-revised}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Vigil (Lady Suyin)} --- Gray-robed, a green lamp at her hip. She never blinks when witnessing an oath.
  \hfill \textit{``The cord will tighten,'' she says. ``I do not need to watch. The river watches for me.''} [\textbf{JUSTICE}] [\textbf{SOCIAL}]
\item \textbf{Cord-Maker (Tama)} --- A woman with hands stained brown from river mud. She weaves silk and hair.
  \hfill \textit{``Each cord is a promise,'' she says. ``The color tells the weight. White for truth. Red for blood. Black for death.''} [\textbf{CRAFT}] [\textbf{TRUTH}]
\item \textbf{Lamp-Keeper (Hiro)} --- An old man with ash on his face. He tends the green flame.
  \hfill \textit{``If the flame goes out,'' he says, ``the river will forget every oath. I cannot sleep.''} [\textbf{RITUAL}] [\textbf{COMMUNITY}]
\item \textbf{River-Spirit (Benzaiten's Messenger)} --- A woman in wet robes, her hair dripping riverweed.
  \hfill \textit{She speaks only in the voices of those who have broken oaths. ``Remember your promise,'' she whispers.} [\textbf{SPIRIT}] [\textbf{TRUTH}]
\item \textbf{Merchant of the Ninth Bale} --- A silk trader who always weighs his bales short by one.
  \hfill \textit{He has never been caught. His witness cord is tied by a Vigil he has bribed.} [\textbf{TRADE}] [\textbf{SOCIAL}]
\item \textbf{Captain of the Monsoon Ship} --- A woman with a brass earring and a scarred hand.
  \hfill \textit{She has never broken an oath. Her witness cord has nine knots. All are tight.} [\textbf{TRAVEL}] [\textbf{HONOR}]
\item \textbf{Oath-Broken Exile} --- A woman with a cut cord on her wrist. She sits apart from the wharf.
  \hfill \textit{``I cut a false cord,'' she says. ``I thought the merchant was honest. He was not. Now I wait.''} [\textbf{COMMUNITY}] [\textbf{REDEMPTION}]
\item \textbf{God-King's Astrologer} --- A man who reads the monsoon and advises on oath timing.
  \hfill \textit{``Never swear an oath during an eclipse,'' he says. ``The river cannot see you.''} [\textbf{POLITICS}] [\textbf{WEATHER}]
\item \textbf{Lamp-Child (Apprentice)} --- A young girl learning to tend the green flame.
  \hfill \textit{She has never spoken. The lamp is her voice. The flame flickers when she is happy.} [\textbf{RITUAL}] [\textbf{INNOCENCE}]
\item[J] \textbf{The Cut Cord's Ghost} --- A translucent strand that floats through the archive at night.
  \hfill \textit{It follows the oath-breaker. It whispers the words of the broken promise.} [\textbf{DEATH}] [\textbf{FEAR}]
\item[Q] \textbf{River-Spirit's Rival (Inari's Fox)} --- A fox with nine tails, seen at the cord-maker's house.
  \hfill \textit{It offers a deal: a cord that cannot break, in exchange for a memory. The memory must be sweet.} [\textbf{MAGIC}] [\textbf{TRICKERY}]
\item[K] \textbf{The First Vigil (Legend)} --- A skeleton in gray robes, sitting at the bottom of the Well of Broken Oaths.
  \hfill \textit{She holds a cord with nine knots. The ninth knot is loose. Untie it, and she will speak.} [\textbf{LEGEND}] [\textbf{DEATH}]
\item[A] \textbf{Benzaiten (The River of Words)} --- A voice heard from the step-pyramid's pool.
  \hfill \textit{``Every word is a current,'' it says. ``Speak gently, or you will drown.''} [\textbf{PATRON}] [\textbf{TRUTH}]
\end{enumerate}

\paragraph*{(Flood/Rumor/Curse)} The river floods --- a bridge is swept away; a rival merchant spreads rumors of cursed silk; the nat of the well is offended --- no one may draw water until appeased.

\section*{Clubs --- Complications/Threats}
\label{sec:lethai-ar-complications-revised}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{River Floods (Bridge Swept)} --- The monsoon rises. The bridge is gone. The wharf is isolated. [\textbf{WEATHER}] [\textbf{TRAVEL}]
  \hfill \textit{The flood is not natural. A river-spirit is angry. An oath was broken upstream.}
\item \textbf{Rumor of Cursed Silk} --- A merchant claims that a bale of silk is cursed. Trade stops. [\textbf{SOCIAL}] [\textbf{TRADE}]
  \hfill \textit{The curse is a lie. The merchant wants to drive down prices. He will buy cheap.}
\item \textbf{Well-Spirit Offended} --- A traveler threw trash into the sacred well. The spirit refuses water. [\textbf{SPIRIT}] [\textbf{DEBT}]
  \hfill \textit{The spirit demands an apology. The apology must be an oath. The oath must be witnessed.}
\item \textbf{Witness Cord Cut (False)} --- Someone cuts a witness cord that was tied in good faith. [\textbf{JUSTICE}] [\textbf{SOCIAL}]
  \hfill \textit{The cutter is the Vigil's own apprentice. She thought the oath was false. It was true.}
\item \textbf{Green Lamp Gutters} --- The eternal flame flickers. The Lamp-Keeper is dying. [\textbf{RITUAL}] [\textbf{COMMUNITY}]
  \hfill \textit{He has one day to train a successor. The successor is the party's least likely candidate.}
\item \textbf{Cord-Maker's Silk Poisoned} --- Someone dyes the silk with a solvent that weakens it. [\textbf{CRAFT}] [\textbf{TRAP}]
  \hfill \textit{The cords will break under pressure. The poisoner is a rival cord-maker.}
\item \textbf{River-Spirit Demands a Story} --- Benzaiten's messenger appears. She asks for a broken oath. [\textbf{SPIRIT}] [\textbf{TRUTH}]
  \hfill \textit{She will not leave until someone confesses. The confession must be true.}
\item \textbf{God-King's False Oath} --- The king swears a vow he does not intend to keep. The Vigil knows. [\textbf{POLITICS}] [\textbf{JUSTICE}]
  \hfill \textit{If she exposes him, the king will exile the Lethai-ar. If she stays silent, the cord tightens on her wrist.}
\item \textbf{Oath-Broken's Redemption Attempt} --- The exile tries to rejoin the Vigil. She offers a true cord. [\textbf{COMMUNITY}] [\textbf{SOCIAL}]
  \hfill \textit{The cord is her own hair. Cutting it would kill her. The Vigil must decide.}
\item[J] \textbf{The Ninth Cup Poured (By Mistake)} --- A stranger pours the ninth cup at a tea house. The Hollow comes. [\textbf{DEATH}] [\textbf{FEAR}]
  \hfill \textit{The Hollow asks: ``Who poured this cup?'' The stranger points at the party.}
\item[Q] \textbf{First Cord Found (Forgery)} --- Someone claims to have found the mythic First Cord. It is a fake. [\textbf{LEGEND}] [\textbf{TRAP}]
  \hfill \textit{The forger is the God-King's astrologer. He wants to control all oaths.}
\item[K] \textbf{River Changes Course} --- The river shifts overnight. The wharf is now on dry land. [\textbf{WEATHER}] [\textbf{DISASTER}]
  \hfill \textit{The river was offended by a broken oath. It will return when the oath is fulfilled.}
\item[A] \textbf{Benzaiten Speaks (Judgment)} --- The river itself speaks from the step-pyramid pool. [\textbf{PATRON}] [\textbf{DEATH}]
  \hfill \textit{``Nine oaths have been broken this season. The ninth was yours. Keep your word or drown.''}
\end{enumerate}

\paragraph*{(Lamp/Cord/Mirror)} Vigil lantern (social edge); silk witness cord (enforce any oath); monsoon mirror (reflects hidden intentions, one use).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:lethai-ar-rewards-revised}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Vigil Lantern} --- A small oil lamp with a green flame. Gain +1 die to social rolls in any Lethai-ar court. [\textbf{SOCIAL}]
  \hfill \textit{The flame flickers when someone lies. It also flickers when you lie.}
\item \textbf{Silk Witness Cord (Crimson)} --- A braided cord dyed with river mud. Use it to enforce any oath. [\textbf{SOCIAL}] [\textbf{JUSTICE}]
  \hfill \textit{The cord tightens when the oath is broken. It will not stop until blood is drawn or the promise is kept.}
\item \textbf{Monsoon Mirror} --- A polished brass disc. Reflects hidden intentions for one scene. [\textbf{TRUTH}] [\textbf{SOCIAL}]
  \hfill \textit{Hold it before a speaker. If they lie, the mirror fogs. The fog smells of river water.}
\item \textbf{Cord-Maker's Needle (Bone)} --- A needle carved from a river-fish's spine. Tie a cord without cutting the silk. [\textbf{CRAFT}]
  \hfill \textit{The needle can also cut a cord without making a sound. Use once.}
\item \textbf{Lamp-Keeper's Ash} --- A pouch of ash from the green flame's hearth. Throw it to calm a river-spirit. [\textbf{SPIRIT}]
  \hfill \textit{The ash is warm. It glows faintly. The spirit will recognize it.}
\item \textbf{River-Spirit's Tear (Vial)} --- A vial of water from the step-pyramid pool. Drink it to recall a forgotten oath. [\textbf{MEMORY}] [\textbf{TRUTH}]
  \hfill \textit{The water tastes of iron. The memory is painful. The pain is the price.}
\item \textbf{Oath-Broken's Cord (Cut)} --- A severed witness cord. Present it to demand a second chance from any Vigil. [\textbf{SOCIAL}] [\textbf{REDEMPTION}]
  \hfill \textit{The cord will re-tie itself if the demand is just. The sound is a whisper.}
\item \textbf{God-King's Witness Seal} --- A golden seal. Commands a Lethai-ar to witness any oath, regardless of price. [\textbf{AUTHORITY}] [\textbf{SOCIAL}]
  \hfill \textit{The seal is heavy. It was stolen from the king's chamber. He will want it back.}
\item \textbf{Bridge of the Ninth Clause Map} --- A chart showing every bridge where contracts can be witnessed. [\textbf{TRAVEL}] [\textbf{LAW}]
  \hfill \textit{The map is written on silk. The silk is a witness cord. Do not cut it.}
\item[J] \textbf{First Cord Fragment (Silk Strand)} --- A piece of the mythic First Cord. Touch it to witness any oath without a cord. [\textbf{LEGEND}] [\textbf{JUSTICE}]
  \hfill \textit{The fragment burns after use. The oath is still binding. The river will remember.}
\item[Q] \textbf{Benzaiten's Whisper (Shell)} --- A river shell that contains the voice of the river-spirit. [\textbf{TRUTH}] [\textbf{SPIRIT}]
  \hfill \textit{Once, ask the river one question. The answer will be true. The river will remember the favor.}
\item[K] \textbf{The Cut Cord's Ghost (Bottle)} --- A ghost of a broken oath sealed in a glass jar. [\textbf{DEATH}] [\textbf{SOCIAL}]
  \hfill \textit{Release it to haunt an oath-breaker. The ghost will whisper the broken promise until the oath is fulfilled.}
\item[A] \textbf{The River's Pardon} --- A scroll signed by Benzaiten herself. Voids any oath, any witness cord. [\textbf{LEGEND}] [\textbf{JUSTICE}]
  \hfill \textit{The scroll is wet. It will never dry. The river will ask for a new oath in exchange.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:lethai-ar-quick-use-revised}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (lanterns, cords, mirrors) that change Position rather than call for a roll.
\item If any A appears, echo \textbf{river-omens}---water that flows upstream, a lamp that burns green without oil, a cord that ties itself.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:lethai-ar-mechanics-revised}

\begin{itemize}
\item \textbf{Witness Clause:} Once per session, when you are present for an oath or contract, you may declare yourself a witness. That oath gains +1 DV to break, and if broken, the oath-breaker must grant you a personal favor (in addition to any other penalty). The witness cord glows faintly for the duration. \textit{Mechanical Note: Player gains 1 Boon if they witness a truthful oath.}
\item \textbf{Lantern of Attendance:} You carry a small oil lamp (the Vigil lantern). As long as it burns, you cannot be surprised in social situations (the flame flickers when deception begins). If the lamp goes out, you lose this benefit for the scene. Refilling the lamp requires a minor offering (a prayer to Benzaiten, a coin in a river). \textit{Mechanical Note: Gain +1 Position on social rolls involving deception detection.}
\item \textbf{Silk Token:} Once per arc, you may offer a witness cord as a token to a faction or NPC. While they hold it, you gain +1 die to all social rolls against them, but they may also call on you to witness one of their own oaths (refusing costs the token and 1 Fatigue). \textit{Mechanical Note: Treat as a String asset.}
\item \textbf{The Ninth Cup Taboo:} In Ayokha, the ninth cup of tea or sake is never poured. Pouring it invites the Hollow to witness the conversation. The Lethai-ar enforce this taboo strictly; a character who pours the ninth cup gains the Hollow's Attention (advance the timer by 1 segment) and must succeed on a Spirit + Resolve test (DV 3) or forget the next minute of conversation. \textit{Mechanical Note: Banks 1 Story Beat for the GM.}
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Benzaiten} --- Aligned with \textit{The Witness}. \textit{The river of words; every oath flows to her.}
    \item \textbf{Inari} --- Aligned with \textit{The Traveler}. \textit{The fox's shadow; contracts are paths, and paths must be witnessed.}
    \item \textbf{The Green Flame} --- Aligned with \textit{The Sealed Gate}. \textit{The lamp that never goes out; the threshold between truth and lie.}
    \item \textbf{The Ninth Current} --- Aligned with \textit{Khemesh}. \textit{The depth beneath the words; the weight of unspoken promises.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{Witness Cords and the Covenant (Mykkiel):} The Covenant recognizes Lethai-ar witness cords as legally binding. A Vigil may be called to testify in a Covenant court; her cord is admissible as evidence.
  \item \textbf{Silk Cords and the Sidhi (Way of Silk):} Sidhi freedmen sometimes carry Lethai-ar witness cords as proof of emancipation. A cord from a respected Vigil is worth more than a freedman's seal.
  \item \textbf{The River Spirit and the Ukwe (Sekogo):} Benzaiten's river is said to flow from the Ukwe's roots. A Lethai-ar who visits the Ukwe can read the water's memory without a context key.
  \item \textbf{The Ninth Cup and the Hollow (Eastern Realms):} The Hollow's Oath-Keepers are attracted to the ninth cup. Aelaerem hedge-keepers never pour a ninth cup in a Lethai-ar's presence; they pour the ninth cup into the ground instead.
\end{itemize}

\subsection*{Lethai-ar Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the River:} Crumble a crust into the current. Ignore the first \emph{River Floods} penalty. [\textbf{SUPERSTITION}]
  \item \textbf{Knot the Cord:} Tie a single knot in a witness cord before using it. Cancel \emph{Witness Cord Cut} or take -1 die on your next oath (your choice). [\textbf{SUPERSTITION}]
  \item \textbf{Name to the Lamp:} Whisper a dead Vigil's name into the green flame. Gain +1 Effect when witnessing an oath this scene, but mark \emph{Obligation} to Benzaiten. [\textbf{SUPERSTITION}]
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:lethai-ar-sb-revised}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Vigils or river-spirits}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{lamp-keepers or cord-makers}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{oath-breakers or the Hollow}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or standing with the Vigil}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local Vigil or wharf-master} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{rising river or lantern failure}.
\item \textbf{Foe Appears:} An \textbf{oath-breaker's ghost or a rival cord-maker} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Vigils, river-spirits, or Hollow agents} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes---\textbf{cord cut/lamp out/river changes course}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Vigil Lady Suyin, River-Spirit, or Benzaiten's voice} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically---\textbf{wharf floods/temple cracks/cords tighten everywhere}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice---\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Lethai-ar (Witness Law):} Green flames flicker in unison, cords glow, a river's current reverses.
\item \textbf{Lethai-ar (Cord Law):} A witness cord tightens without cause, a knot unties itself, a cut cord re-knots.
\item \textbf{Lethai-ar (River Law):} Water levels rise without rain, a fish speaks a name, the ninth cup pours itself.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth):} On a hit, shift one band. On a strong hit, also impose \emph{Lantern Glare}.
\item \textbf{Cut the Cord (Presence + Sway):} On a hit, freeze range for one exchange; if a witness cord is in your possession, +1 die.
\item \textbf{Weather the Flood (Body + Athletics):} On a hit in \emph{River Floods}, you pick who reaches high ground; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Cord \& River)}
\begin{itemize}
\item \textbf{The Cord-Cutter (Nine Broken Oaths)} --- \emph{Phases}: Witness Cords Severed $\rightarrow$ Orchestrated Perjury $\rightarrow$ Vigil Coup. \emph{Cracks}: re-tie a cut cord, expose a false Vigil, survive a night in the Well of Broken Oaths.
\item \textbf{The River-Dryer (Benzaiten's Wrath)} --- \emph{Phases}: Water Levels Dropping $\rightarrow$ Procession of Empty Wharves $\rightarrow$ Flood-Baptism of a Merchant. \emph{Cracks}: relic river water, survivor testimony, the green flame burning without oil.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Lethai-ar of Ayokha, power flows through witness cords, lantern flames, and the memory of the river. A Vigil's authority rests on the number of oaths she has witnessed and never broken. The lamp is the symbol of that authority; a Vigil without a lamp is not a Vigil.

\textbf{For the GM:}  
Power in Lethai-ar society is measured in cords tied, lamps kept lit, and oaths witnessed. Patronage often takes the form of lanterns, witness cords, or mirrors --- each tied to a specific Vigil, wharf, or contract. To emphasize the cultural mechanics:
\begin{itemize}
\item Use the \textbf{Witness Clause} to create social dilemmas: should the party witness an oath that might be broken?
\item Enforce the \textbf{Lantern of Attendance} --- a party member who carries a Vigil lantern gains social advantages but must maintain the flame.
\item Let \textbf{Silk Token} be a form of communal currency: a witness cord given to a faction creates a mutual obligation.
\item Track the \textbf{Ninth Cup Taboo} as a source of Hollow Attention. A single mistake can summon an Oath-Keeper.
\end{itemize}
In Lethai-ar society, the river forgets nothing --- and the witness cord remembers every word.
\end{tcolorbox}

% Index entries
\index{Lethai-ar (Ayokha)|see{River-Bound Oath-Keepers}}
\index{Theme generator}
\index{River-Bound Oath-Keepers generator}
\index{Places!Lethai-ar}
\index{People!Lethai-ar}
\index{Complications!Lethai-ar}
\index{Rewards!Lethai-ar}
\index{Witness cords}
\index{Vigil lantern}
\index{Silk token}
\index{Ninth cup taboo}
\index{Benzaiten}
\index{River spirits}
\index{Monsoon wharves}
\index{Oath-Broken exile}
\index{Local Patrons!Lethai-ar}