\section{Lethai−al −−− ``The Storm−Touched Wanderers of Nihori''}
łabel{chap:lethai−al}

\textbf{Key Demographics: } Lethai−al
\textbf{Capital City: } Morikage

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The road is a prayer. The storm is our law. The ninth wave remembers your name.}

\begin{quote}
  \textbf{Procession Elder Kenji:}  
  ``The sea does not forgive, but it does forget. If you can wait long enough, the waves will loosen what binds you. But we do not wait. We walk. From shrine to shrine, from storm to storm, we carry the names of the drowned. The road is a prayer. Every step is a syllable.''
\end{quote}

\begin{quote}
  \textbf{Tide−Charm Keeper Yuki:}  
  ``My grandmother drowned in the ninth wave. Her name became a tide−charm. I keep it in a pouch of oiled leather. When the storm grows angry, I whisper her name. The sea remembers her. It calms for a moment −−− long enough to breathe. We are wanderers. The road is our home. The dead travel with us.''
\end{quote}

\textbf{Genre / Mood:} Processional pilgrimage, wandering monks, duelists at crossroads, storm−folk, ancestral debt.

\textbf{Starting Location:} A procession trail winding along the sea cliffs, marked by weathered torii gates and small stone shrines. A group of Lethai−al walk in single file, chanting a counting song. One carries a palanquin with a covered object −−− a tide−charm shrine. The leader stops: ``You stand on the processional road. The drowned are watching from the waves. Do you walk with us? If so, you must carry nothing. The road provides. The road takes away.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme & Atmosphere}]
The Lethai−al of Nihori once guarded a drowned kingdom −−− an archipelago that sank when a forbidden name was spoken. Their blood still carries the salt of that ancient catastrophe. Now they are wanderers in the oldest sense: masterless, landless, bound to no lord but fiercely bound to each other and to the storm spirits that swirl above the waves. They walk from shrine to shrine, from port to port, following the processional roads that circle the island. They do not farm. They do not build. They carry everything they own on their backs or in wooden carts. Their duels are not to the death −−− they are to the first draw of blood, after which the loser must join the procession and carry the winner's burden for a season. Their courts are crossroads. Their temples are the road itself. The Eastern Ninth lives here: the Hollow does not hide under hills −−− it breathes in the salt foam. The ninth wave is never the largest; it is the one that remembers your name.

\vspace{2mm}
\textbf{What you notice first:} The smell of salt and incense. The way the Lethai−al never sit with their backs to the sea. The constant, low murmur of waves being predicted −−− not by instruments, but by the tilt of a head, the twitch of a finger. The worn wooden plaques hung from their belts, each carved with a dead ancestor's name.

\textbf{The one rule that kills newcomers:} Never walk under a torii gate without making an offering. The gate is a threshold. The road on the other side is not the same road you left.
\end{tcolorbox}

\subsection*{Regional Mood: The Processional Road}

Power here is measured in shrines visited, storms survived, and the number of dead ancestors carried as tide−charms. The Lethai−al have no king, no courts, no written law. Their authority is personal and processional: a Procession Elder's word is law because she has walked the circuit more times than anyone else. Their guilds are not guilds −−− they are processional bands, each with its own rhythm, its own counting song, its own patron shrine. A band that walks too fast offends the spirits; one that walks too slow loses the tide. The Lethai−al do not believe in ownership of land; they believe in passage rights, storm oaths, and the memory of the drowned. Every crossroads is a court. Every storm is a messenger. Every sunrise is a departure.

\textbf{Starting Location (alternative):} A crossroads shrine at the edge of a fishing village. A procession of Lethai−al has stopped for the night. Their carts are arranged in a circle, painted with wave patterns. A Procession Elder sits by a small fire, offering cups of sake to travelers who pause. ``The road is long. The dead are many. Stay one night. Tell us a story. If we like it, we will carry it to the next shrine. If we do not, you will walk alone tomorrow.''

\subsection*{Prominent NPCs of the Lethai−al}
łabel{sec:lethai−al−npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
ℎline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
ℎline
Procession Elder Kenji & Leader of the Ninth Wave Band & Complete the circuit before the monsoon & He has walked the circuit 99 times. He plans to stop at 100. \\
ℎline
Tide−Charm Keeper Yuki & Guardian of drowned names & Prevent the Hollow from taking her grandmother's charm & She carries an oiled pouch with nine charms. The ninth is empty −−− for herself. \\
ℎline
Tide−Hunter Rona & Leader of the Harpoon Guild & Kill the sea monsters before they rise & She rides the monsters rather than killing them; she knows their names. \\
ℎline
The Hollow−Bound & A corrupted ancestor spirit & Reclaim the names stolen by the living & It speaks through the mouths of those who use tide−charms. \\
ℎline
The Drowned Captain (Ghost) & A Lethai−al who died before completing the circuit & Ask the living to carry his charm to the final shrine & He appears on foggy nights, holding a tide−charm that hums. \\
ℎline
\end{longtable}

\subsection*{Names of the Lethai−al}
łabel{sec:lethai−al−names}

Inspired by Japanese and coastal Ainu traditions: short, with vowel endings, honorifics (−san, −sama, −dono). Surnames denote a band, a wave, or a drowned ancestor.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
ℎline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Band/Clan} & \textbf{Tide−Name} \\
ℎline
Kenji & Yuki & Ninth Wave & \textit{of the Drowned Shore} \\
ℎline
Ronan & Hana & Foam−kin & \textit{of the Salt Breath} \\
ℎline
Mason & Mira & Storm−walker & \textit{of the Empty Pouch} \\
ℎline
Takeshi & Sora & Shore−bearer & \textit{of the Ancestor's Shell} \\
ℎline
Jiro & Nami & Circuit−keeper & \textit{of the Unfinished Road} \\
ℎline
\end{longtable}

\paragraph*{(Crossroads/Grove/Shrine)} A crossroads shrine with a missing stone; a bamboo grove where duels are fought at dawn; a processional torii gate at the edge of a cliff.

\section*{Spades −−− Places}
łabel{sec:lethai−al−places}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem \spadecard{1} \textbf{Crossroads Shrine (Stone)} −−− A small stone shrine at the meeting of three roads.
  ℎfill \textit{Older travelers leave a pinch of salt on the altar. The spirits take it. The salt never depletes.} \texttt{[THRESHOLD]}
ℹtem \spadecard{2} \textbf{Bamboo Grove (Duel Ground)} −−− A clearing where the stalks are scarred by blade cuts.
  ℎfill \textit{The bamboo bleeds sap when a duel is dishonorable. The sap is red.} \texttt{[DUEL]}
ℹtem \spadecard{3} \textbf{Torii Gate (Processional)} −−− A wooden gate standing alone on a cliff path.
  ℎfill \textit{Walk through it, and the wind changes direction. The gate was built by a drowned mason.} \texttt{[THRESHOLD]}
ℹtem \spadecard{4} \textbf{Cart Circle (Night Camp)} −−− A ring of painted wooden carts, their wheels touching.
  ℎfill \textit{Inside the circle, the Hollow cannot enter. Outside, the road waits.} \texttt{[COMMUNITY]}
ℹtem \spadecard{5} \textbf{Tide−Charm Vault (Sea Cave)} −−− A cave where processional bands store charms during the winter.
  ℎfill \textit{The charms hum when the sea is angry. The cave is watched by a blind keeper.} \texttt{[MEMORY]}
ℹtem \spadecard{6} \textbf{Storm Shrine (Peak)} −−− A stone circle with bells that ring in the wind.
  ℎfill \textit{Each bell has a different tone. The tones predict the weather. The shrine maiden reads them.} \texttt{[WEATHER]}
ℹtem \spadecard{7} \textbf{Foam−Ford (Crossing)} −−− A shallow channel that can only be crossed at low tide.
  ℎfill \textit{The crossing lasts nine breaths. Count carefully. The ninth breath belongs to the sea.} \texttt{[TRAVEL]}
ℹtem \spadecard{8} \textbf{Volcano Shrine (Dormant)} −−− A crater where the name that sank the kingdom was spoken.
  ℎfill \textit{The crater is warm. Speaking any name aloud here causes the ground to tremble.} \texttt{[MYSTERY]}
ℹtem \spadecard{9} \textbf{Sake House of the Ninth Cup} −−− A tavern at a crossroads, where duels are negotiated.
  ℎfill \textit{The ninth cup is never poured. It is for the dead. The dead sometimes drink it.} \texttt{[SOCIAL]}
ℹtem \spadecard{J} \textbf{Ancestor's Salt Pan} −−− A flat where seawater is evaporated to make salt.
  ℎfill \textit{The salt is mixed with ash from funeral pyres. It tastes of grief. The widows sell it to remember.} \texttt{[LABOR]}
ℹtem \spadecard{Q} \textbf{The Sunken Gate} −−− An underwater archway that leads to the drowned kingdom.
  ℎfill \textit{It can only be reached by holding your breath for nine heartbeats. The gate is guarded by a spirit.} \texttt{[DEATH]}
ℹtem \spadecard{K} \textbf{Procession Elder's Lookout} −−− A wooden tower on the highest cliff.
  ℎfill \textit{From here, the Elder can see three storms coming. He has never been wrong.} \texttt{[WEATHER]}
ℹtem \spadecard{A} \textbf{The Drowned Capital (Myth)} −−− A sunken city said to rise once every century.
  ℎfill \textit{Those who enter must share a breath with the dead. The dead are talkative.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Elder/Keeper/Ronin)} Procession Elder Kenji with a tally stick; Tide−Charm Keeper Yuki with an oiled pouch; Wandering Ronin with a sake cup.

\section*{Hearts −−− People & Factions}
łabel{sec:lethai−al−people}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem ℎeartcard{1} \textbf{Procession Elder} −−− Gray−haired, sunburned, a tally stick in his belt.
  ℎfill \textit{``The circuit is 108 shrines. I have walked it 99 times. The 100th will be my last.''} \texttt{[AUTHORITY]}
ℹtem ℎeartcard{2} \textbf{Tide−Charm Keeper} −−− A young woman with a leather pouch at her hip. She never swims.
  ℎfill \textit{``My grandmother is in this pouch. She cannot get wet. When the storm comes, I whisper her name.''} \texttt{[MEMORY]}
ℹtem ℎeartcard{3} \textbf{Tide−Hunter} −−− A woman who hunts sea monsters. She has a harpoon made from a ship's keel.
  ℎfill \textit{She has never killed a monster. She has ridden nine of them. They recognize her whistle.} \texttt{[COMBAT]}
ℹtem ℎeartcard{4} \textbf{Wandering Ronin} −−− A man with no master, a scarred blade, and a cup of sake.
  ℎfill \textit{He offers a cup to every opponent. ``Drink. The duel will be fair. The loser carries the winner's burden for a season.''} \texttt{[DUEL]}
ℹtem ℎeartcard{5} \textbf{Shrine Maiden} −−− A woman with a copper bell on a rope. She rings it at dawn.
  ℎfill \textit{The bell's tone tells the procession whether to walk or wait. It has never been wrong.} \texttt{[RITUAL]}
ℹtem ℎeartcard{6} \textbf{Cart Builder} −−− A silent craftsman who builds the painted carts. He uses no nails.
  ℎfill \textit{``Nails are for coffins. The road does not forgive splinters.'' The joints are held by counting songs.} \texttt{[CRAFT]}
ℹtem ℎeartcard{7} \textbf{Wave−Walker} −−− A woman who can walk on water during a storm. She does not know how.
  ℎfill \textit{``The storm carries me. I just don't fight it. The road is wet, but the road is still the road.''} \texttt{[MYSTIC]}
ℹtem ℎeartcard{8} \textbf{Sake Brewer} −−− A fat man with a wooden ladle. He brews sake from rice and seawater.
  ℎfill \textit{``The salt makes it strong. The sea makes it honest. The dead like it. It helps them remember.''} \texttt{[TRADE]}
ℹtem ℎeartcard{9} \textbf{The Hollow−Bound} −−− A corrupted ancestor spirit speaking through a living host.
  ℎfill \textit{It demands a name. It offers power. It takes memory.} \texttt{[THREAT]}
ℹtem ℎeartcard{J} \textbf{Drowned Scholar (Ghost)} −−− A translucent figure who haunts the Sunken Gate.
  ℎfill \textit{He asks: ``Has anyone spoken the name?'' He will not say which name. He has been asking for a century.} \texttt{[DEATH]}
ℹtem ℎeartcard{Q} \textbf{Volcano Priest} −−− A man who tends the dormant crater. He speaks only in whispers.
  ℎfill \textit{``The name that sank the kingdom is still here. Do not ask what it is. The crater knows. It is waiting.''} \texttt{[MYSTIC]}
ℹtem ℎeartcard{K} \textbf{The Ninth Wave's Herald} −−− A child who appears before every storm. She points at the horizon.
  ℎfill \textit{She never speaks. She only points. The direction tells where the storm will hit. She has never been wrong.} \texttt{[OMEN]}
ℹtem ℎeartcard{A} \textbf{The First Drowned (Legend)} −−− The first Lethai−al who died in the sinking. She sits at the bottom of the sea.
  ℎfill \textit{If you bring her a memory, she will teach you to breathe water. The memory must be your own.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Duel/Pirate/Ashfall)} A duel challenge is issued −−− refusal loses face; a pirate flotilla claims the strait −−− pay toll or fight; ashfall from the volcano −−− all actions Desperate.

\section*{Clubs −−− Complications/Threats}
łabel{sec:lethai−al−complications}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem \clubcard{1} \textbf{Duel Challenge (Issued)} −−− A Lethai−al challenges a party member to a duel. Refusal loses face. \texttt{[SOCIAL]}
  ℎfill \textit{The duel is to first blood. The loser carries the winner's burden for a season, or shares a story to mend the road.}
ℹtem \clubcard{2} \textbf{Pirate Flotilla} −−− Kalingan pirates block the strait. They will not let you pass without a tide−charm. \texttt{[COMBAT]}
  ℎfill \textit{The charm must be from a drowned ancestor. Refuse, and they attack. The pirates count in nines.}
ℹtem \clubcard{3} \textbf{Ashfall (Volcano)} −−− The dormant crater emits ash. Visibility drops. All actions are Desperate. \texttt{[WEATHER]}
  ℎfill \textit{The ash is warm. It smells of old names. Speaking a name makes the ash thicken.}
ℹtem \clubcard{4} \textbf{Storm−Spirit Offended} −−− A traveler insulted the wind. The storm−spirit demands amends. \texttt{[OATH]}
  ℎfill \textit{What it wants is a promise about the sea. Break it, and the spirit drowns you.}
ℹtem \clubcard{5} \textbf{Tide−Charm Stolen} −−− Someone steals a pouch of charms. The ancestors are angry. \texttt{[THEFT]}
  ℎfill \textit{The thief is a foreign merchant. The charms are in a cargo hold. The ship sails at dawn.}
ℹtem \clubcard{6} \textbf{Drowned Capital Rising} −−− The sunken city emerges at low tide. Spirits walk the streets. \texttt{[DEATH]}
  ℎfill \textit{They ask for news of the surface. Lies will anger them. Truths will sadden them. Both are dangerous.}
ℹtem \clubcard{7} \textbf{Volcano Name Whispered} −−− Someone speaks the forbidden name near the crater. The ground shakes. \texttt{[DISASTER]}
  ℎfill \textit{The speaker is a child. The child did not know. The volcano remembers. It will not forget.}
ℹtem \clubcard{8} \textbf{Ninth Wave Warning} −−− The full moon is tomorrow. The ninth wave will be larger than any in a century. \texttt{[WEATHER]}
  ℎfill \textit{The fishing fleet cannot be recalled in time. Someone must warn them. The wave is already forming.}
ℹtem \clubcard{9} \textbf{Sake House Brawl} −−− A dispute over a duel's fairness turns violent. The ninth cup is poured. \texttt{[COMBAT]}
  ℎfill \textit{The ninth cup is for the dead. Someone has just died. The brawl is now a blood feud.}
ℹtem \clubcard{J} \textbf{Sunken Gate Opens} −−− The underwater archway becomes passable. Something swims out. \texttt{[DEATH]}
  ℎfill \textit{It is a drowned spirit. It wants its name back. It carries a secret. Give it what it wants, and you will learn the truth.}
ℹtem \clubcard{Q} \textbf{Procession Elder's Silence} −−− The Elder stops reading the weather. The band does not know where to walk. \texttt{[CRISIS]}
  ℎfill \textit{He is grieving. His son was lost at sea. He will not speak until the body is found.}
ℹtem \clubcard{K} \textbf{Duel to the Death (Betrayal)} −−− A duel that was meant to be to first blood becomes a killing. \texttt{[COMBAT]}
  ℎfill \textit{The killer is a spy from Ashaan. The duel was an assassination. The party is blamed.}
ℹtem \clubcard{A} \textbf{The Ninth Wave Takes} −−− The massive wave crashes over the cliff shrine. The bells are silent. \texttt{[APOCALYPSE]}
  ℎfill \textit{The shrine maiden is gone. The storm−spirit is free. The Hollow walks the shore.}
\end{enumerate}

\paragraph*{(Charm/Arrow/Cup)} Storm−charm (one weather bypass); broken arrow (proof of oath); sake cup (right to demand a fair duel).

\section*{Diamonds −−− Rewards/Leverage}
łabel{sec:lethai−al−rewards}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem ⋄card{1} \textbf{Storm−Charm} −−− A bone carved with wave patterns. Use it to ignore one weather penalty on a journey leg. \texttt{[WEATHER]}
  ℎfill \textit{The bone is from a sailor who died in a storm. His spirit protects you. He asks for a story to carry.}
ℹtem ⋄card{2} \textbf{Broken Arrow (Oath−Proof)} −−− A snapped arrow shaft. Present it to prove an oath was fulfilled. \texttt{[SOCIAL]}
  ℎfill \textit{The arrow was broken by a duelist who lost fairly. It carries his honor. It is warm to the touch.}
ℹtem ⋄card{3} \textbf{Sake Cup (Duel Right)} −−− A ceramic cup used in a fair duel. Present it to demand a duel to first blood. \texttt{[SOCIAL]}
  ℎfill \textit{The cup is chipped. The chip is from a blade that drew first blood. The loser bears the burden.}
ℹtem ⋄card{4} \textbf{Tide−Charm (Leather Pouch)} −−− A small pouch containing a drowned ancestor's name on a shell. \texttt{[MEMORY]}
  ℎfill \textit{Whisper the name to calm a storm for one hour. The ancestor will remember. \textbf{Tick Hollow Timer +1}.}
ℹtem ⋄card{5} \textbf{Procession Pass (Wooden Plaque)} −−− A carved name−tag. Grants safe passage with any Lethai−al band. \texttt{[TRAVEL]}
  ℎfill \textit{The name is a dead ancestor. The band will recognize it. They will feed you.}
ℹtem ⋄card{6} \textbf{Shrine Maiden's Bell (Copper)} −−− A copper bell. Ring it to pacify a storm−spirit for one scene. \texttt{[RITUAL]}
  ℎfill \textit{The bell's tone is unique. The spirit will recognize it. It may ask for another promise.}
ℹtem ⋄card{7} \textbf{Volcano Priest's Whisper} −−− A single word: the name that is forbidden. Speak it to cause a tremor. \texttt{[MAGIC]}
  ℎfill \textit{The word burns the tongue. Use it once. The volcano will remember. The ground will crack.}
ℹtem ⋄card{8} \textbf{Wave−Walker's Sandal} −−− A single sandal woven from sea grass. Wear it to walk on calm water for one minute. \texttt{[TRAVEL]}
  ℎfill \textit{The sandal will not get wet. Your foot will. The water is cold.}
ℹtem ⋄card{9} \textbf{Drowned Captain's Oar} −−− A skeletal oar that weighs nothing. Use it to row any boat against the current. \texttt{[TRAVEL]}
  ℎfill \textit{The captain's ghost will row with you. He will ask for a story. He knows if it is true.}
ℹtem ⋄card{J} \textbf{Ninth Wave's Shell} −−− A white shell found on the Ninth Wave's Rock. It contains a memory. \texttt{[MEMORY]}
  ℎfill \textit{Break it to experience the memory. The memory is always a death. The death is always a drowning.}
ℹtem ⋄card{Q} \textbf{Procession Elder's Flute} −−− A bone flute. Play it to predict the weather for the next 24 hours. \texttt{[WEATHER]}
  ℎfill \textit{The tune is sad. It always predicts a storm. It is always right. The storm will find you.}
ℹtem ⋄card{K} \textbf{Sunken Gate Key (Coral)} −−− A piece of coral shaped like a key. Opens the Sunken Gate. \texttt{[DEATH]}
  ℎfill \textit{The key is alive. It grows. It will only fit the gate once. The gate will remember your face.}
ℹtem ⋄card{A} \textbf{First Drowned's Breath (Glass Vial)} −−− A vial of seawater from the bottom of the sunken capital. \texttt{[LEGEND]}
  ℎfill \textit{Drink it to breathe water for one hour. The taste is salt and sorrow. Your lungs will ache.}
\end{enumerate}

\section*{Quick use notes}
łabel{sec:lethai−al−quick−use}
\begin{itemize}
ℹtem Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2−−5 $→$ 4, 6−−10 $→$ 6, J/Q/K $→$ 8, A $→$ 10).
ℹtem Diamonds are codified outcomes (charms, arrows, cups) that change position rather than call for a roll.
ℹtem If any A appears, echo \textbf{processional omens}−−−a gate that creaks without wind, a cart wheel that turns backward, a name that echoes from the sea.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
łabel{sec:lethai−al−mechanics}

\begin{itemize}
ℹtem \textbf{Processional Circuit:} The Lethai−al walk a circuit of 108 shrines. A character who walks the circuit (over several sessions) may, at the end, gain +1 die to all rolls involving storms, spirits, or the sea. The circuit takes nine months. A character who completes the circuit nine times becomes a Procession Elder.
ℹtem \textbf{Walking Court (Crossroads Justice):} When two Lethai−al dispute, they settle it at a crossroads. Each party speaks their truth once. Then they walk away in opposite directions. The road decides. A character who lies at a crossroads gains the condition \textit{Wave−Walker's Shame} (−1 die to all social rolls until they walk the circuit again).
ℹtem \textbf{Storm−Reading:} Once per journey, a character with Lethai−al training may predict the weather for the next 24 hours (Wits + Survival vs. DV 3). On success, choose one leg to ignore weather penalties. On a miss, the storm finds you anyway, and the GM gains 1 Story Beat to spend on a thunderclap or flash flood.
ℹtem \textbf{Tidal Resilience:} Once per scene, when a character would take Harm from a water−based or storm−based effect, they may reduce it by 1 level (to a minimum of 0). Mark 1 Fatigue.
ℹtem \textbf{The Eastern Ninth (Hollow in Salt Foam):} When a Lethai−al drowns, their name becomes a tide−charm. Living kin may whisper it to calm a storm, but the Hollow remembers. At the next full moon, the Hollow draws closer. The Hollow's attention ticks whenever a tide−charm is used.
\end{itemize}

\subsection*{Lore Echoes (Cross−Expansion Hooks)}
\begin{itemize}
  ℹtem \textbf{Tide−Charms and the Tulkani (Way of Silk):} Tulkani story−tellers sometimes carry Lethai−al tide−charms as proof of a witnessed drowning. A charm can be traded for a story that the Tulkani have not heard. The story must be about the sea.
  ℹtem \textbf{Storm−Spirits and the Veiled Sisters (Ashaan):} The Veiled Sisters have tried to bind storm−spirits for use in their fleet. The Lethai−al have foiled nine attempts. A Procession Elder may ask the party to sabotage a Sister's ritual.
  ℹtem \textbf{The Drowned Capital and the Kalrantha Amulet (Dhahara):} The sunken city is said to contain a temple that holds a fragment of the demon Varash. The amulet's creator visited the capital before it sank. The fragment whispers the forbidden name.
  ℹtem \textbf{The Ninth Wave and the Hollow (Eastern Realms):} The Hollow's Oath−Keepers ride the ninth wave. Lethai−al who survive the ninth wave are said to walk beyond the Hollow's reach. Aelaerem leave offerings of salt at the high tide line to greet the wave.
\end{itemize}

\subsection*{Lethai−al Superstitions (burn for +1 die once)}
\begin{itemize}
  ℹtem \textbf{Bread to the Foam:} Crumble a crust at the water's edge. Ignore the first \emph{Pirate Flotilla} penalty. \texttt{[SUPERSTITION]}
  ℹtem \textbf{Knot the Rope:} Tie a single knot in a shrine climbing rope. Cancel \emph{Storm−Spirit Offended} or take −−1 die on your next sailing roll (your choice). \texttt{[SUPERSTITION]}
  ℹtem \textbf{Name to the Wave:} Whisper a drowned ancestor's name into the ninth wave. Gain +1 Effect when navigating a storm this scene, but mark 1 Fatigue. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
łabel{sec:lethai−al−sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
ℹtem \textbf{Exposure:} Your actions draw unwanted attention from \textbf{storm−spirits or pirate scouts}.
ℹtem \textbf{Noise:} Sounds of your actions alert nearby \textbf{shrine maidens or duelists}.
ℹtem \textbf{Trace:} Evidence of your passage marks your route for \textbf{tide−hunters or the Hollow}.
ℹtem \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}.
ℹtem \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
ℹtem \textbf{Alarm Raised:} \textbf{Local Procession Elder or shrine maiden} becomes aware and begins responding.
ℹtem \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{ashfall or rising tide}.
ℹtem \textbf{Foe Appears:} A \textbf{pirate crew or a storm−spirit} arrives on scene.
ℹtem \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
ℹtem \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
ℹtem \textbf{Reinforcements:} Additional \textbf{pirates, storm−spirits, or volcanic ash clouds} arrive.
ℹtem \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
ℹtem \textbf{Major Twist:} The situation fundamentally changes−−−\textbf{duel becomes death/storm shifts/sunken gate opens}.
ℹtem \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
ℹtem \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
ℹtem \textbf{Trap Springs:} A prepared danger activates with full effect.
ℹtem \textbf{Authority Arrival:} \textbf{Procession Elder Kenji, Drowned Captain, or Ninth Wave Herald} intervenes.
ℹtem \textbf{Scene Shift:} The environment changes dramatically−−−\textbf{volcano erupts/storm breaks/sunken capital rises}.
ℹtem \textbf{Patron Omen:} Divine/arcane forces take notice−−−\textbf{omen appears/blessing lost/curse manifests}.
ℹtem \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region−Specific SB Options}
\begin{itemize}
ℹtem \textbf{Lethai−al (Processional Law):} A torii gate creaks, a cart wheel turns backward, a name echoes from the sea.
ℹtem \textbf{Lethai−al (Duel Law):} Blades glow faintly, spilled sake steams, the first drop of blood evaporates.
ℹtem \textbf{Lethai−al (Drowned Law):} A name appears in wet sand, a wave carries a message, a drowned face looks up from still water.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
ℹtem \textbf{Advance/Withdraw (Wits + Sail):} On a hit, shift one band. On a strong hit, also impose \emph{Storm Cover}.
ℹtem \textbf{Cut the Duel (Presence + Sway):} On a hit, freeze range for one exchange; if a sake cup is in your possession, +1 die.
ℹtem \textbf{Weather the Storm (Resolve + Survival):} On a hit in \emph{Storm−Spirit Offended}, you pick who is struck by lightning; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Wave & Procession)}
\begin{itemize}
ℹtem \textbf{The Name−Eater (Nine Drownings)} −−− \emph{Phases}: Tide−Charms Stolen $→$ Orchestrated Storm $→$ Drowned Coup. \emph{Cracks}: recover a stolen charm, expose a false ancestor, survive the ninth wave.
ℹtem \textbf{The Unforgotten Storm (Spirit of the Ninth Wave)} −−− \emph{Phases}: Waves Rising $→$ Procession of Empty Shrines $→$ Foam−Baptism of a Storm−Reader. \emph{Cracks}: relic storm−charm, survivor testimony, the ninth wave breaking on a calm sea.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage & Power}]
Among the Lethai−al, power flows through the procession, the road, and the memory of the drowned. A Procession Elder's word is law because he has walked the circuit more than anyone else. A Tide−Charm Keeper holds the names of the dead; she can calm a storm or call one. There are no kings, no courts, no written laws −−− only the road, and the road does not negotiate.

\textbf{For the GM:}  
Power in Lethai−al society is measured in shrines visited, storms survived, and tide−charms kept. Patronage often takes the form of storm−charms, broken arrows, or processional passes −−− each marked by a specific ancestor, duel, or shrine. To emphasize the cultural mechanics:
\begin{itemize}
ℹtem Use \textbf{Processional Circuit} as a long−term character goal. Completing it is a major achievement.
ℹtem Enforce \textbf{Walking Court} −−− a character who lies at a crossroads suffers social penalties until they make amends.
ℹtem Let \textbf{Storm−Reading} be a journey−saving ability, but with a risk of additional complications.
ℹtem Track the \textbf{Hollow's Attention} when tide−charms are used. The Oath−Keepers may appear at the worst moment.
\end{itemize}
In Lethai−al society, the sea forgets nothing −−− and the ninth wave remembers every name.
\end{tcolorbox}

% Index entries
∈dex{Lethai−al|see{Storm−Touched Wanderers}}
∈dex{Theme generator}
∈dex{Storm−Touched Wanderers generator}
∈dex{Places!Lethai−al}
∈dex{People!Lethai−al}
∈dex{Complications!Lethai−al}
∈dex{Rewards!Lethai−al}
∈dex{Processional circuit}
∈dex{Walking court}
∈dex{Storm−reading}
∈dex{Tide−charms}
∈dex{Drowned kingdom}
∈dex{Ninth wave}
∈dex{Hollow in salt foam}
∈dex{Duel customs}
∈dex{Storm−spirits}
∈dex{Sunken capital}
∈dex{Local Patrons!Lethai−al}