\section{Linn --- ``The Oath-Bound''}
\label{chap:linn}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The sea writes in salt. The frost binds in iron. The skald remembers in verse.}

\begin{quote}
  \textbf{High Jarl (Elite):}  
  ``The sea keeps better counsel than any scribe, and the storm remembers every oath sworn upon its winds. But the frost --- the frost is the witness. It preserves what the tide would wash away. A promise sworn in winter is owed until the ice melts. And the ice never truly melts.''
\end{quote}

\begin{quote}
  \textbf{Net-Wife (Commoner):}  
  ``The water speaks to those who know how to listen --- the way it moves tells you if the fish are running, if the weather's turning, or if something's coming that wasn't invited to the feast. My grandmother taught me to read the frost on the oar-blades. That's how I know who will live through the winter and who will feed the seals.''
\end{quote}

\textbf{Genre / Mood:} Norse coastal saga, oath-driven honor, horror of the unwitnessed death, the tyranny of seasons, the weight of wyrd.

\textbf{Starting Location:} A longship house lined with storm-charms and whale-bone runes, salt wind clawing at the doors. A Net-Wife binds a storm-token into a PC's palm. ``The Whale-Road sings,'' she says. ``But the Mistlands whistle back. Bring us three keel-lengths of ironwood from the Valewood edge, or the tide will take our boats in the spring. The frost remembers the shelter you took last winter. Pay, or the ice will hold it against you.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
North of the Aberderrin Sea, where fjords split the coast and skerries break the tide, dwell the Linn. They are traders and raiders in equal breath, their longships crowding the whale-road and river mouths. They descend the Ylolka like an axe-blade and cut into southern shores, or slip through the Thin Strip where the Valewood dares not touch salt. Yet even these storm-oath people dread the Mistlands and the Valewood, whose fog and trees have no need of ships or harbors to claim their due.

The Linn measure their lives in seasons, not years. There is a season for raiding (summer, when the seas are open), a season for planting (spring, when the frost breaks), a season for governance (autumn, when the ships are beached and the Thing meets), and a season for survival (winter, when the frost sets its mark). In winter, the fjords freeze, the longships are dragged into sheds, and the people gather around the longhall fires to settle debts, tell sagas, and wait for the sun to return. The winter debt is not a book --- it is a set of carved oath-sticks, each notch a promise kept or broken, each knot a story still unfolding. When a Linn says a debt is ``written in frost,'' they mean it is unchangeable until the ice melts. And the ice, in the high north, never truly melts.

The gods of the Linn are not distant. \textbf{Adar}, the All-Father, walks the ridges in a gray cloak, his one eye watching for those who break oaths. He is a manifestation of the Oath of Flame and Light, but the Linn know him as the god of wisdom, death, and binding promises. \textbf{Rorg}, the Storm Father, rides the clouds, his hammer cracking the sky. He is the patron of raiders and sailors, a Thor-like figure who laughs in the teeth of the gale. His cult is loud, proud, and dangerous --- a Rorg devotee will swear a storm-oath by drinking seawater and shouting at the lightning. The Temple of Light calls him a demon. The Linn call him captain.

\vspace{2mm}
\textbf{What you notice first:} The smell of tar and salt fish. The creak of oars from unseen boats. The way everyone --- jarl to thrall --- touches a knotted cord before speaking of the sea. The oath-sticks hanging from every belt, each mark a weight carried.

\textbf{The one rule that kills newcomers:} Never name a drowned person aloud while underway. The sea collects names, and if it hears you speaking one it doesn't recognize, it may reach up to correct the debt. Also, never refuse a cup of winter mead --- it is a test of your honor and your weight.
\end{tcolorbox}

\subsection*{The Almanac of Ordinary Days}
(A commoner's view, overheard at a winter hearth)

\textit{What do people eat for breakfast?} Salt fish and sour bread --- the fish is smoked over peat fires, the bread is leavened with fog-damp yeast. The yeast is caught in cloth traps hung from the bell-towers. The traps are emptied at dawn. The yeast is not alive. It is not dead either.

\textit{What does a child learn before they can walk?} To ring a bell. Every Mistlands child is given a small bronze bell at birth. By age three, they can ring the eight notes of the ward-chime in their sleep. The ninth note is never taught. The ninth note is discovered. Discovery is fatal.

\textit{What is the first thing you notice about a stranger?} Their salt-pouch. Every Mistlander carries a leather pouch of consecrated salt. A stranger without a pouch is either a fool or a wraith. A stranger whose pouch is empty is a fool who is about to become a wraith.

\textit{What is the most common crime?} Ward-salt theft. The Aeler control the salt pans, and the salt is the only thing that keeps the dead from rising. A stolen pouch can buy a week's protection. The price is a year's weight to the Aeler. The weight is collected in labor. The labor is collecting salt. The salt is stolen. The cycle is perpetual.

\textit{What does no one talk about but everyone knows?} The Bell-Wardens are not chosen --- they are volunteered. A village that loses a warden must send a replacement to the Aeler training holds. The holds are in the mountains. The mountains are cold. The trainees do not always return. The ones who do return do not speak of what they learned. Their bells ring at a different pitch.

\textit{Where do people go when they die?} To the fog. Bodies are not buried in the Mistlands --- the ground rejects them. Corpses are carried to the edge of the bell-line and left for the mist. The mist takes them. Sometimes, at dusk, you can see them walking the lanes. They do not speak. They do not need to. Their presence is a question.

\textit{How do you apologize for a serious wrong?} You ring the Bell of Atonement --- a cracked bronze bell that hangs in every village square. The bell rings once for each wrong confessed. If the bell rings true, you are forgiven. If it cracks further, you are not. If it shatters, you are banished to the fog. The banished never return. Their voices still answer from the white.

\textit{What is the most important festival?} The Rehanging of the Lines (autumn). Every bell-line is taken down, cleaned, and rehung. The work takes three days. The villages are silent for three nights. The dead listen. The dead wait. The dead do not ring.

\textit{What is the secret that could destroy a powerful person?} The Legate of the Mists is a wraith. She died in 803 AR, during a bell-line failure. The Aeler rebuilt her from fog and memory. She does not know she is dead. She attends council meetings. She signs documents. Her signature fades after sunset. The documents are refiled every morning.

\subsection*{Wyrd and the Tv\'{\i}h\"{o}gg --- The Double-Cut Fate}
\label{sec:linn-wyrd}

The Linn do not believe in a single, woven fate. They believe in \textbf{wyrd} --- a web of actions and consequences that can be \textit{cut}, not rewoven. A skilled volva (seeress) or a desperate warrior may invoke the \textbf{tv\'{\i}h\"{o}gg} (``double-cut''), severing one thread of fate to redirect another.

Mechanically, once per session, a character who has sworn a storm-oath or carries a frost mark may declare a \textbf{tv\'{\i}h\"{o}gg} before rolling. They choose one:
\begin{itemize}
    \item \textbf{Cut the Weight:} Treat a failed roll as a success, but the GM gains 1 SB and the character marks a \textbf{Frost Mark} (as if they had accepted a gift --- the weight follows them).
    \item \textbf{Cut the Witness:} Reroll any roll involving the sea, storms, or oaths, but the character's name is added to the ice in a hidden place. The GM may call on this weight once in the future (a complication, a recalled debt, a spirit asking for a name).
\end{itemize}
The tv\'{\i}h\"{o}gg cannot be undone. The Linn say: \textit{``A cut thread does not grow back. You simply learn to carry the shorter end.''}

\subsection*{Regional Mood: Skerry Stormcall, Frost Memory}

Power here is spoken, not written. A jarl's authority rests on the weight of his oath-chest and the sharpness of his skald's tongue. Longships are not vessels; they are promises carved in timber. Every voyage is a gamble against tide, reef, and the attention of things that sleep beneath the waves. The Linn do not fear death --- they fear dying without a witness, swallowed by fog where no verse will remember their name.

But beneath the sea-lore runs the frost-bound memory. Each Linn keeps an oath-stick, notched for every weight they carry and every debt owed to them. When the winter gales ground the longships, the Thing convenes to settle the winter debts. A jarl whose stick is too deeply notched may find his hall burned; a thrall who has collected enough marks may buy his freedom. What is carved in ice is the true law of the Linn --- not the spoken word, but the carved one.

\subsection*{Prominent NPCs of Linn}
\label{sec:linn-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Jarl Skjold & High Jarl of the Southern Fjords & Unite the scattered strands under one oath & His beard is braided with silver from every ship he has taken. His oath-stick bears notches from three wars. \\
\hline
Volva Hervor & Mist-seer and oath-reader & Read the tides of fate before they turn & She speaks once a year; the rest of the time she listens to the fog. She reads the frost-bound oaths by touch --- her fingers are scarred from counting. \\
\hline
Astrid & Net-Wife (commoner exemplar) & Keep her village safe from the Hollow and the deep & She ties a knot for every storm she's survived. Her belt has thirty. She is the keeper of the village oath-sticks. \\
\hline
Brother Halvard & Missionary of the Temple of Light & Convert the Linn to the Lens of the White Flame & He carries a water-clock to measure the hours of prayer, but the skerries have no hours. His oath-stick is blank --- the Linn do not trust him. \\
\hline
Gunnhild & Sea-Queen's Hand & Enforce the queen's will without drawing a blade & She never draws her sword. The threat is enough. She carries an oath-stick notched with the names of every jarl who tried to cheat the queen. \\
\hline
The Draugr Jarl & Unfinished business (rumored) & A dead jarl whose oath was never fulfilled & Said to rise from the Thing-Holm when the ninth bell tolls. His oath-stick is frozen --- the marks are filled with ice. \\
\hline
Einar Frost-Tongue & Skald of the Frost Memory & Keep the memory of weights alive through verse & He has never lost a verse duel. His tongue is forked --- he can praise and curse in the same breath. His oath-stick is longer than his arm. \\
\hline
\end{longtable}

\subsection*{Names of Linn}
\label{sec:linn-names}

Inspired by the northern fjords: short, hard consonants, compound words that evoke the sea and the storm. Surnames are often patronymic (-son, -dottir) or taken from a notable deed.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Skjold & Astrid & Einarsson & \textit{the Oath-Keeper} \\
\hline
Bjorn & Gunnhild & Hervorsdottir & \textit{Tide-Reader} \\
\hline
Ragnar & Sigrid & Ketilsson & \textit{Bell-Sworn} \\
\hline
Halvard & Yrsa & Ormsson & \textit{the Unfinished} \\
\hline
Torsten & Runa & Sturladottir & \textit{Whale-Bone} \\
\hline
Vidar & Helga & Haakonsson & \textit{Ash-Prow} \\
\hline
Einar & --- & Frost-Tongue & \textit{the Remembered} \\
\hline
\end{longtable}

\paragraph*{(Fjord/Skerry/Hall)} Kelp-skerry gut: a narrow channel where oars scrape weed and rock; tide-shed of longship houses with tar smoke and gulls for sentries.

\section{Spades --- Places}
\label{sec:linn-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Kelp-Skerry Gut} --- Narrow channel where oars scrape weed and rock.
  \hfill \textit{Kelp strands like dead men's hair; row silently or the reef listens.} [TRAVEL] [STEALTH]
\item \textbf{Tide-Shed} --- Longship houses; tar smoke and gulls for sentries.
  \hfill \textit{Each shed holds a keel's ghost; old crews sleep under the floor planks.} [REPUTATION]
\item \textbf{Wave-Gate Reef} --- Just off a raiding beach; only locals know the cut.
  \hfill \textit{One wrong turn and the wave breaks over you --- twice.} [TRAVEL] [DEATH]
\item \textbf{Runestone Causey} --- Across tidal flat; names half lost to barnacle.
  \hfill \textit{Touch the stone with your left hand. If it's warm, the ancestors approve.} [LEGEND]
\item \textbf{Herring-Stairs} --- Drying racks above cliff path; knives flicker like fish.
  \hfill \textit{Workers sing a counting song --- eight verses, never nine.} [LABOR]
\item \textbf{Winter Hall} --- Turf walls, smoke-eye roof, benches crowded with shields.
  \hfill \textit{The fire never goes out. Feed it a story each night or it goes hungry.} [COMMUNITY]
\item \textbf{Mistlands Maze} --- Reed-channels braid and vanish; navigation by instinct.
  \hfill \textit{The reeds whisper names you've forgotten. Do not answer.} [DEATH] [FEAR]
\item \textbf{Boomed Harbor} --- Iron rings in whale-bone posts; controlled access.
  \hfill \textit{The chain was forged from a broken oath. It rattles when liars approach.} [AUTHORITY]
\item \textbf{Aberderrin Headland} --- Current rips past wreck timbers in the kelp.
  \hfill \textit{On foggy nights, the drowned signal with lanterns. Never follow.} [DEATH]
\item[J] \textbf{Dolmis Waystation} --- Low isle with fresh water and old weights.
  \hfill \textit{The well is lined with copper. A single dropped coin returns as a blessing.} [TRAVEL]
\item[Q] \textbf{Thing-Holm} --- Flat islet ringed with stones for speaking and swearing.
  \hfill \textit{The stones have ears; oaths spoken here echo for three tides. The winter oaths are kept here, frozen in a block of ice that never melts.} [JUSTICE]
\item[K] \textbf{High Jarl's Seat} --- Above fjord, beacon tower glaring out to sea.
  \hfill \textit{The beacon oil is mixed with whale fat. It burns green when a storm is close, blue when the winter debts are called.} [WEATHER]
\item[A] \textbf{Whale-Road Horizon} --- Open water where storms are born and names are made.
  \hfill \textit{Here, the sky meets the sea. Here, the dead meet the living.} [LEGEND]
\end{enumerate}

\paragraph*{(Jarl/Oar-Crew/Keeper)} Net-wife who reads the sea like a memory and the sky like a mood; steersman whose tiller-hand never shakes.

\section{Hearts --- People \& Factions}
\label{sec:linn-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Net-Wife} --- Reads sea like memory, sky like mood; weather-witch in all but name.
  \hfill \textit{She ties a knot for every storm she's survived. Her belt has thirty. She is the keeper of the village oath-sticks.} [COMMUNITY]
\item \textbf{Steersman} --- Tiller-hand never shakes, even when keel does.
  \hfill \textit{His left ear is scarred --- he listened too long to a siren's promise. His oath-stick is notched for every life he's lost.} [REPUTATION]
\item \textbf{Shipwright} --- Pitch on palms, schedule in teeth; knows every plank's story.
  \hfill \textit{He names each timber after a dead relative. The keel is his grandmother. The ship's burden is carved into the mast.} [CRAFT]
\item \textbf{Oar-Master} --- Sets beat that decides if home is today or never.
  \hfill \textit{He drums on a hollowed skull. The rhythm is a prayer to Rorg. The oar-crew's debt is marked on his own arm --- notches in the bone.} [SOCIAL]
\item \textbf{Mist-Pilot} --- Ghosts boats through marsh lanes; reed-runner with secret paths.
  \hfill \textit{He never blinks in fog. Some say he traded his blink for safe passage. His memory-cord is knotted, not notched --- the knots are spells.} [STEALTH]
\item \textbf{Skald} --- Verses buy silence, start fights, end them; law in rhyme.
  \hfill \textit{Her tongue is forked --- she can praise and curse in the same breath. She is the only one who can reshape a frost-bound oath --- by composing a new verse that changes the meaning of a mark.} [JUSTICE] [SOCIAL]
\item \textbf{Shield-Band} --- Cousins fine with coin, better with cattle.
  \hfill \textit{They sharpen their swords on runestones. The sparks tell fortunes. Their oath-stick is shared --- each mark is a blood debt owed to the band.} [COMBAT]
\item \textbf{Foster-Son} --- Southern coast birth; speaks both laws and subtext.
  \hfill \textit{He wears two brooches --- one iron, one bone. Neither is his true allegiance. He carries two oath-sticks --- one for the Linn, one for the lowlands.} [POLITICS]
\item \textbf{Thing-Speaker} --- Makes war into lawsuit, lawsuit into lunch.
  \hfill \textit{He carries a water-clock that runs on meltwater. When it runs out, so does your patience. He interprets the winter debts for those who cannot read ice.} [JUSTICE]
\item[J] \textbf{Sea-Queen's Hand} --- Envoy with seal-bag and sword-belt; queen's will made manifest.
  \hfill \textit{He never draws his blade. The threat is enough. His oath-stick is branded with the queen's mark --- a seal that cannot be forged.} [AUTHORITY]
\item[Q] \textbf{Sea-Queen} --- Crown of hammered silver, eyes for weather and law.
  \hfill \textit{She lost an eye to a storm --- or to a spy. She won't say which. She is the only one who can order the winter debts spoken.} [AUTHORITY]
\item[K] \textbf{High Jarl} --- First among raids, last to boast; paramount over strands.
  \hfill \textit{His beard is braided with silver from every ship he's taken. His oath-stick is a whale's rib --- too long for anyone else to carry.} [REPUTATION]
\item[A] \textbf{Volva of Mist} --- Oath-reader whose nod stills surf; seer of tides and truth.
  \hfill \textit{She speaks once a year. The rest of the time, she listens to the fog. She reads the frost-bound oaths by touch --- her fingers are scarred from counting.} [LEGEND]
\end{enumerate}

\paragraph*{(Weather/Reef/Feud)} Black squall drops from clear band; oars or prayers --- pick one; fogfall makes friend and reef trade places.

\section{Clubs --- Complications/Threats}
\label{sec:linn-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Black Squall} --- Drops from clear band; oars or prayers --- pick one. [WEATHER]
  \hfill \textit{The wind screams a name. Listen carefully --- it might be yours. Rorg is testing your courage.}
\item \textbf{Fogfall} --- Sound travels, sight does not; friend and reef trade places. [WEATHER] [FEAR]
  \hfill \textit{You hear oars. Then you don't. Then they're behind you. The drowned are rowing.}
\item \textbf{Boom Lift} --- Harbor chain rises; schedules drown in harbor mouth. [AUTHORITY]
  \hfill \textit{The chain's links are inscribed with names of those who carry weight. Each one rattles differently. Your name is on one.}
\item \textbf{Levy Clash} --- Viterra's wardens arrive with polite spears. [POLITICS] [COMBAT]
  \hfill \textit{They carry a writ stamped with the queen's seal. The seal is still wet. The winter ice says the queen is owed three ships.}
\item \textbf{Keel-Rot Rumor} --- Crews beach mid-voyage; trust in timber fails. [DISASTER]
  \hfill \textit{The rot smells sweet --- too sweet. Someone poisoned the tar. The ship's mark is missing.}
\item \textbf{Feud Token} --- Raid becomes wedding... or funeral; oath-debt called due. [SOCIAL] [DEBT]
  \hfill \textit{A bronze ring arrives on an arrow. The meaning is clear. The ice shows the mark.}
\item \textbf{Oath Matured} --- Verse sworn last winter comes due today. [SOCIAL] [DEBT]
  \hfill \textit{The skald clears her throat. The room goes cold. The ice in the Thing-Holm cracks.}
\item \textbf{Mistlands Miscount} --- Channels ``move,'' cargo vanishes without witnesses. [DEATH]
  \hfill \textit{The fog lifts to reveal a different shore. Your charts are now lies. The cargo's mark is blank --- someone erased it.}
\item \textbf{Aberderrin Race} --- Contrary currents split flotilla; half late, half lost. [WEATHER]
  \hfill \textit{The water churns like a dying man. Something beneath is feeding. The sea remembers a new debt.}
\item[J] \textbf{Southron Fire-Pots} --- Night-run turns to daylight; surprise ashore. [COMBAT]
  \hfill \textit{Pots shatter on the deck. The fire burns blue --- alchemical. Rorg laughs. The raid's memory is now fire-scarred.}
\item[Q] \textbf{Thing Injunction} --- No sailing till grievance heard (hours? days?). [JUSTICE]
  \hfill \textit{The Thing-Speaker smiles. He has all the time in the world. The winter oath is frozen --- no one can leave until it thaws.}
\item[K] \textbf{General Muster} --- Horns on every headland; all bays become spears. [AUTHORITY] [COMBAT]
  \hfill \textit{The war fleet assembles. Your ship is the smallest. Adar is watching. He counts the oaths.}
\item[A] \textbf{Ground-Sea} --- Long, blind swell rolls under; steering lies. [WEATHER] [DEATH]
  \hfill \textit{The swell has a rhythm --- two short, one long. A signal from below. The drowned are voting.}
\end{enumerate}

\paragraph*{(Right/Token/Escort)} Harbor-green mark --- dock without levy in named strand; oar-share claims two benches on fast longship.

\section{Diamonds --- Rewards/Leverage}
\label{sec:linn-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Harbor-Mark} --- Dock without levy in named strand (once). [TRAVEL]
  \hfill \textit{The mark is a notch cut in a bone. Match it to the harbor post. The notch must be cut by the harbor master's own knife.}
\item \textbf{Oar-Share} --- Claim two benches on a fast longship for a season. [TRAVEL]
  \hfill \textit{The oars are carved with your name. Row well, or they'll be recarved. The oar itself is the promise.}
\item \textbf{Pilot's Token} --- Mist-runner guides through reed-maze safely. [TRAVEL]
  \hfill \textit{The token is a dried eel skin. Keep it dry. The pilot's memory-cord is knotted in the same pattern.}
\item \textbf{Wharf-Right} --- Dolmis waystation loading priority; first ashore. [TRAVEL]
  \hfill \textit{The harbormaster frowns but steps aside. Your name is now on a list.}
\item \textbf{Salvage Claim} --- Flotsam on reef stretch becomes ``inventory.'' [TRADE]
  \hfill \textit{You must name each piece before touching it. The wrong word forfeits the claim. The promise of the salvage is bound to the tide line.}
\item \textbf{Thing Ruling} --- Verdict in your favor that travels with you. [JUSTICE]
  \hfill \textit{The verdict is carved on a whale-rib. Carry it as a shield. The winter ice will not speak for the same dispute twice.}
\item \textbf{Raid-Truce} --- Immunity in named bay for one market day. [TRAVEL] [SOCIAL]
  \hfill \textit{A white banner on your mast. No one will attack --- openly. The promise of the truce is the banner rope. Cut it, and the truce ends.}
\item \textbf{Foster-Bond} --- Guest-kin by jarl's house; hospitality guaranteed. [SOCIAL]
  \hfill \textit{You share a cup with the jarl. The cup is silver. The promise is iron. The winter ice bears a new mark --- for both of you.}
\item \textbf{Herring Allotment} --- Priority barrels when the run hits; silver in season. [TRADE]
  \hfill \textit{The barrels are sealed with your mark. Break one to test quality. The promise of the allotment is the barrel hoop.}
\item[J] \textbf{Escort Writ} --- Eight shields and wolf-banner from cape to cape. [TRAVEL] [COMBAT]
  \hfill \textit{The shields are painted with a wolf's head. They die before you do. Their oath is their swords --- notched for every oath kept.}
\item[Q] \textbf{Queen's Audience} --- Whisper exits as policy; sea-law bends. [AUTHORITY]
  \hfill \textit{The Queen's hand writes your words on vellum. The ink is her blood. The winter ice will record the audience --- but not the words.}
\item[K] \textbf{High Jarl's Pennon} --- Levies open booms and mouths when it flies. [AUTHORITY]
  \hfill \textit{The pennon is torn from a captured ship. Fly it proudly --- or else. The promise of the pennon is the ship's mast.}
\item[A] \textbf{Storm-Oath Clause} --- Temporary exception to any sea-law, sealed. [AUTHORITY] [MAGIC]
  \hfill \textit{The clause is whispered, never written. The sea hears and obeys --- once. Rorg will remember. The winter ice will not record it --- the ice cannot hold a whisper.}
\end{enumerate}

\section{Quick use notes}
\label{sec:linn-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (rights/tokens/escorts) that change position rather than call for a roll.
\item If any A appears, echo \textbf{sea-omens} --- horns on wind, white horses on swell, verse unfinished until landfall, the ice cracking with memory.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:linn-mechanics}

\begin{itemize}
\item \textbf{The Frost-Burden (Weight Clock):} Every Linn community keeps a winter rod or a frozen block of ice carved with marks representing promises and weights. A character who accepts a favor, shelter, or gift from a Linn must accept a \textbf{Frost Mark} --- a 4-segment weight clock. When the clock fills, the community may demand a service (a raid, a story, a season of labor). A character who refuses to accept a Frost Mark is considered an oath-breaker and suffers $-1$ die on all social rolls with Linn for the remainder of the journey. The Frost Mark can be cleared by performing a season-appropriate task (planting in spring, raiding in summer, governing in autumn, surviving in winter).
\item \textbf{Storm-Oaths:} A Linn may swear an oath by touching salt water. Breaking a storm-oath inflicts 1 Fatigue (the sea's disappointment) and gives the GM a free SB to represent the sea's attention. Keeping an oath until its completion grants $+1$ die on the next roll involving the sea or sailing. Oaths sworn during a storm (under Rorg's eye) are doubly binding --- breaking them inflicts 1 Harm.
\item \textbf{The Season's Turn:} The GM may declare that the current season affects certain actions. In \textbf{Spring} (planting), foraging and healing rolls gain $+1$ die. In \textbf{Summer} (raiding), combat and sailing rolls gain $+1$ die. In \textbf{Autumn} (governance), social and legal rolls gain $+1$ die. In \textbf{Winter} (survival), all survival rolls are Desperate, but clearing a Frost Mark is easier (the community is too cold to argue). The season changes when the GM spends 3 SB, or after a major story beat.
\item \textbf{Saga Reputation (Community Track):} Every Linn settlement has a Saga Track [6] that reflects the community's oral memory. The party can increase it by performing deeds worthy of verse (rescuing a ship, slaying a sea monster, keeping a difficult oath). At 3 segments, locals offer hospitality without question. At 6, the party gains a permanent epithet and may call on one favor from any Linn jarl.
\item \textbf{Mist-Dread:} Linn characters suffer $-1$ die on navigation rolls when entering the Mistlands or Valewood, but gain $+1$ die on rolls to avoid supernatural fear (they have learned to dread the right things).
\item \textbf{Skald's Burden:} A PC who publicly composes a verse about an event may choose to increase or decrease the Saga Track by 1 (as they shape memory). On a failed Performance roll (DV 3), the verse backfires and the track moves the opposite direction.
\item \textbf{The Wyrd Track (Optional):} Each PC may maintain a personal Wyrd Track [6]. The GM ticks it when the character invokes a tv\'{\i}h\"{o}gg, breaks a small taboo, or is named in a skald's verse. At 3 ticks, the character feels the weight of unseen attention --- the first time each session they roll a 1, the GM may convert it into a complication of their choice. At 6 ticks, the character is \textbf{fated}: once, they may treat a failed roll as a success, but the GM gains 3 SB and the character's name is added to a hidden winter mark somewhere in the Linn lands. The Wyrd Track resets only after a major sacrifice or a pilgrimage to a volva's cairn.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Adar the All-Father} --- Oath of Flame and Light. \textit{The one-eyed god of wisdom, death, and binding oaths. He watches from the ridges. His oath-stick is the frost-bound memory itself.}
    \item \textbf{The Storm Father} --- Rorg. \textit{The god of thunder, raiding, and the wild sea. His devotees swear storm-oaths by drinking seawater. He laughs at lightning.}
    \item \textbf{The Salted Anchor} --- The Sealed Gate. \textit{The harbor that never closes, the oath that never rusts.}
    \item \textbf{The Wave Mother} --- Ra\'{e}yn. \textit{The tide that carries and the storm that answers.}
    \item \textbf{The Drowned Chorus} --- Khemesh. \textit{The pressure beneath the keel, the weight the sea collects.}
    \item \textbf{The Unwritten Verse} --- The Witness. \textit{The memory that survives the drowning.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Draugr Jarl (Unfinished Business):} The Thing-Holm is haunted by a jarl who died before his oath was fulfilled. He rises when a storm-oath is broken, demanding a blood-price paid in stories. The Temple of Light has sent three missionaries to lay him to rest; all three drowned in calm water. His oath-stick is frozen into the ice of the Thing-Holm. If it ever thaws, he will walk again.
  \item \textbf{Adar's Ravens (The Witness):} Adar is said to have two ravens, Memory and Thought, who fly across the world each day and return to whisper in his ear. A Linn skald who prays to Adar may ask one question about a forgotten weight; the GM must answer truthfully, though cryptically. The price is a memory --- the skald forgets something unimportant, but it may become important later.
  \item \textbf{Rorg's Hammer (The Xhak'thul):} The Storm Father's hammer is said to be lodged in a cliff on the highest skerry. A character who touches the hammer and survives the lightning may invoke Rorg's wrath once per session: all enemies within Near must test Resolve (DV 4) or be Stunned for one exchange. The cost is that the character is marked by the storm --- they will be struck by lightning at the worst possible moment (GM's discretion).
  \item \textbf{The Tide-Washed Covenant (The Hospice of the Salted Anchor):} A gray-robed healer runs a hospice on the edge of the Mistlands, treating wounded raiders and fog-mad sailors alike. She asks no payment, but she writes every name in a book that never dries. Linn captains leave offerings of fresh water before a voyage; those who refuse find their ship plagued by the drowned. The book is said to be a copy of the frost-bound oaths --- but written in salt water, not ice.
  \item \textbf{The Ninth:} In any longhouse, the ninth horn blown at a feast summons the Hollow. The wise stop at eight. Those who blow the ninth are said to see a pale figure standing at the high table --- waiting to collect a weight. The figure is Adar, checking his oath-stick.
\end{itemize}

\subsection*{Linn Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Wake:} Toss a crust over the gunwale before a voyage. Ignore the first \emph{Black Squall} penalty. [SUPERSTITION]
  \item \textbf{Knot the Oar:} Tie a single knot in your oar's leather grip. Cancel \emph{Keel-Rot Rumor} or take $-1$ die on your next rowing check (your choice). [SUPERSTITION]
  \item \textbf{Name to the Deep:} Whisper a dead enemy's name into the tide. Gain +1 Effect against his kin for the scene, but mark \emph{Obligation} to the Sea-Queen. [SUPERSTITION]
  \item \textbf{Drink to the Storm:} Pour a cup of mead into the sea before a raid. Rorg will grant you favorable winds --- but he will also expect a share of the plunder. [SUPERSTITION]
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:linn-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{coastal guards or rival crews}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{ships or harbor watch}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{trackers or mist-pilots}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}.
\item \textbf{Supply Strain:} Mark $+1$ segment on a relevant \textbf{resource clock}.
\item \textbf{Frost Mark:} The GM may spend 1 SB to add a segment to a character's Frost Mark clock (as the community remembers a forgotten weight).
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local jarl or harbor master} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{weather shift or reef hazard}.
\item \textbf{Foe Appears:} A \textbf{rival crew or levy force} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\item \textbf{Season Turns:} The GM may spend 2 SB to advance the season to the next phase, with associated bonuses/penalties.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{ships, warriors, or skalds} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes --- \textbf{storm hits/oath invoked/feud declared}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\item \textbf{Winter Debts Called:} The Thing is convened. The party must defend their oaths or be forced into a quest.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Sea-Queen, High Jarl, or Volva} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically --- \textbf{storm rises/reef shifts/mist closes in}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice --- \textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\item \textbf{Adar's Judgment:} The All-Father appears (or sends a raven). He demands that a broken oath be mended or a weight paid in full. Refusal costs 2 SB.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Linn (Storm-Oaths):} Sudden calm, breaking waves speak names, verses come unbidden to lips.
\item \textbf{Linn (Mistlands):} Channels shift without sound, voices call from reed-beds, compass spins.
\item \textbf{Linn (Thing-Law):} Witnesses appear, oaths are demanded, skald begins composing your saga.
\item \textbf{Linn (Winter Memory):} Oath-sticks crack, ice sheets groan, a forgotten debt surfaces in a dream.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Sail + Wits):} On a hit, shift one band. On a strong hit, also impose \emph{Wake Surge}.
\item \textbf{Cut the Lane (Pilot + Tactics):} On a hit, freeze range for one exchange; if a beacon is under your color, $+1$ die.
\item \textbf{Weather the Squall (Resolve + Sail):} On a hit in \emph{Black Squall}, you pick who drifts; on a miss, the Keeper does.
\end{itemize}

\subsection*{Boss Seeds (Sea \& Signal)}
\begin{itemize}
\item \textbf{The Net-Father (Nine Nets)} --- \emph{Phases}: Harbor Favors $\rightarrow$ Orchestrated Wrecks $\rightarrow$ Thing Coup. \emph{Cracks}: cut a cord lineage; reveal swapped verdict; survive a staged squall.
\item \textbf{The Blue Cantor (Brine-Gospel)} --- \emph{Phases}: Drownings Called Mercy $\rightarrow$ Procession at Reef $\rightarrow$ Tide-Baptism of a Jarl. \emph{Cracks}: relic earplugs, survivor testimony, storm turning against them.
\item \textbf{The Frost Jarl (Oath-Keeper)} --- \emph{Phases}: Mark Calls $\rightarrow$ Ice Cracks $\rightarrow$ The Thing of Winter. \emph{Cracks}: pay a forgotten wergild, speak a true name from the winter ice, survive a night in the Thing-Holm with the Draugr Jarl.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Linn, power flows through oaths, kinship, and the spoken word --- and through the marks carved in ice. Jarls maintain authority through successful raids and wise judgments, while skalds wield influence through their ability to immortalize deeds in verse --- or damn names through satire. The Sea-Queen holds sway over the coastlands, but her rule depends on the respect of the High Jarl and the Thing-speakers who interpret ancient laws. And beneath all of them lies the frost-bound memory --- the carved word, not the written one.

\textbf{For the GM:}  
Power in Linn society is fluid and performance-based. Patronage often takes the form of foster-bonds, oar-shares, or raid-truces --- each binding the recipient to service and reputation. To emphasize this:
\begin{itemize}
\item Tie rewards to visible symbols (tokens, banners, verses) that can be challenged, stolen, or voided.
\item Let rival jarls issue conflicting claims, forcing players to choose whose favor matters more.
\item Use the Thing-holm and harbor gatherings as arenas for social contests, where oaths are sworn and broken.
\item Introduce the Frost Mark as a recurring pressure. Every gift from a Linn comes with a weight. Players who ignore their marks may find themselves summoned to the Thing in winter.
\item Let the seasons change. A winter scene should feel different from a summer raid. Use the Season's Turn to reinforce the cyclical nature of Linn life.
\end{itemize}
In Linn society, your word is your bond, and your reputation is your greatest treasure. But the frost-bound memory is patient. It waits for spring.
\end{tcolorbox}

\subsection*{Temple Crusade Mini-Generators}

\subsubsection*{Temple Pressure Events (d6)}
Roll when the Northern Crusade advances or when the GM needs a story beat.
\begin{enumerate}[label=\arabic*., wide=0pt]
    \item \textbf{Grain Fleet Arrives.} Temple ships unload cheap southern grain. Prices drop, but the old shrine offerings are forgotten. \textit{Effect:} $+1$ tick on \textbf{Temple Pressure} clock; $+1$d on social rolls with hungry villagers for one scene.
    \item \textbf{Synod Court Opens.} A Thepyrgosi judge hears the first case. Feuds may be settled with penance instead of blood-silver. \textit{Effect:} Shrine-devout characters suffer $-1$ Position in this settlement until they prove their loyalty.
    \item \textbf{Bell Tower Consecrated.} A bronze bell rings nine times at dawn. The old shrine bells fall silent. \textit{Effect:} Temple-leaning characters gain $+1$d on faith actions; every midnight, the GM may spend 1 SB to have the ninth bell attract a hungry ghost.
    \item \textbf{Schoolmaster Arrives.} Children learn to write Temple script. Old oral traditions are called ``superstition.'' \textit{Effect:} The next generation will have $+1$ Temple Devotion; players may attempt a \textbf{Synthesis Scene} to save one shrine teaching.
    \item \textbf{Bishop's Progress.} A high-ranking cleric tours the coast, consecrating altars. Refusal to host him marks you as an enemy of the Light. \textit{Effect:} Begin a \textbf{Bishop's Favor} [4] clock; if it fills, the bishop demands a public oath of loyalty.
    \item \textbf{Iconoclasm.} Zealots burn a seastack shrine or break a Thing's oath-ring. The old patrons are furious. \textit{Effect:} Shrine-devout characters gain a free \textbf{Vengeance Oath} String; weather and tide hazards become one step worse for 1d4 days.
\end{enumerate}

\subsubsection*{Linn Reactions to the Temple (d12)}
How a community (or a crew) responds to the crusade's advance.
\begin{enumerate}[label=\arabic*., wide=0pt]
    \item \textbf{Desperate Acceptance.} ``We ate their grain. We owe them.'' \textit{Tag:} Grateful--Sullen.
    \item \textbf{Silent Resistance.} Shrine offerings are left in secret; Temple priests are given cold hearths. \textit{Tag:} Ice--Hospitality.
    \item \textbf{Open Feud.} A clan declares the Temple an enemy. Raids on granaries begin. \textit{Clock:} \textbf{Feud with Temple} [6].
    \item \textbf{Syncretic Curiosity.} Elders ask the priest to bless a shrine. ``Both fires warm the hall.'' \textit{Opportunity:} Earn a \textbf{Synthesis Token}.
    \item \textbf{Pragmatic Split.} The jarl converts; half the thing follows, half walks out. \textit{Clock:} \textbf{Thing Schism} [6].
    \item \textbf{Conversion as Revenge.} A family spurned by the old patrons embraces the Temple to spite them. \textit{Tag:} Bitter--Righteous.
    \item \textbf{Bread \& Bells.} The village accepts the Temple but keeps a hidden shrine to Hv\'{\i}tabj\"{o}rn. \textit{Tag:} Double--Altar.
    \item \textbf{Heretic Burnings.} Temple zealots execute a shrine-keeper. The hall is now \textbf{Contested Ground}.
    \item \textbf{The Bishop's Bastard.} A priest's secret child is now a Linn raider. \textit{String:} Blood--Secret.
    \item \textbf{Wave-Father's Answer.} A storm wrecks a Temple grain ship. Shrine-devout see a sign. \textit{Effect:} $+2$ on a shrine-related roll this session.
    \item \textbf{Oath-Ring Saved.} A Temple judge rules that the old oath-ring is still lawful -- a rare bridge-builder. \textit{Reward:} \textbf{Ambassador Status} (one scene of neutral faith ground).
    \item \textbf{The Silent Bell.} The new tower's bell cracks on the ninth toll. No one can explain it. \textit{Omen:} GM gains 2 SB for fog- or ghost-related twists.
\end{enumerate}

\subsubsection*{Temple Missions (d6)}
Quests the Temple of Light might offer to a mixed or Temple-leaning crew.
\begin{enumerate}[label=\arabic*., wide=0pt]
    \item \textbf{Recover a Stolen Relic.} A shrine-devout crew has taken a fragment of the True Candle. Find it -- by negotiation or force.
    \item \textbf{Escort the Bishop.} Take a southern prelate through a contested fjord. Protect him from ``pagan'' ambushes -- or arrange one.
    \item \textbf{Convert a Hall.} A jarl sits on the fence. Persuade him (or his skald) that the Light is good for trade. \textit{Clock:} \textbf{Conversion} [6].
    \item \textbf{Burn the Shrine.} A remote seastack shrine is a navigation hazard and a gathering point for resisters. The Temple offers silver for its destruction.
    \item \textbf{Rescue a Synod Judge.} He was taken hostage after a biased ruling. Infiltrate the clan's winter hall and bring him out alive.
    \item \textbf{Forgive a Blood-Debt.} The Temple wants to end a feud by paying blood-silver from its own coffers. You must convince both sides to accept -- a delicate parley.
\end{enumerate}

\subsubsection*{Faith Tension Scenes (d6)}
Quick scene starters for religious conflict.
\begin{enumerate}[label=\arabic*., wide=0pt]
    \item \textbf{The Thinglet at the Hearth.} Shrine-devout and Temple-leaning families argue over a broken betrothal. \textit{Position:} Controlled $\rightarrow$ Desperate if you take sides openly.
    \item \textbf{Midnight at the Bell Tower.} Someone has silenced the ninth bell -- by stuffing it with wool from a white bear's pelt. \textit{Complication:} Spiritual Conflict Condition begins.
    \item \textbf{The Convert's Shame.} A Temple-leaning Linn begs you to smuggle her family heirloom (a shrine totem) out of the village before the bishop's agents find it.
    \item \textbf{Storm over the Synod Court.} V\'{e}lki answers a shrine-devout's prayer with a lightning strike on the Temple granary. \textit{GM:} Spend 2 SB to have a priest accuse the party of witchcraft.
    \item \textbf{The Wandering Skald.} A singer arrives with a verse that claims the crew has a secret shrine oath. The verse is false -- but now everyone believes it.
    \item \textbf{Dual Oath at Dusk.} A dying warrior asks to swear by both Hr{\ae}vkn and the Light. The shrine-keeper and the Temple priest each demand you witness only their rite. \textit{Reward:} A \textbf{Synthesis Token} if you honor both.
\end{enumerate}

\subsubsection*{Syncretic Outcomes (d6)}
When players successfully blend shrine and Temple practices, roll (or choose) a lasting effect.
\begin{enumerate}[label=\arabic*., wide=0pt]
    \item \textbf{The Shared Cairn.} A new shrine uses a Temple candle to light the old oath-ring. \textit{Effect:} This site is now \textbf{Mixed Quarter} permanently; once per session, any character may reroll a failed faith-related roll here.
    \item \textbf{Witness of Both.} The crew is known as bridge-builders. \textit{Effect:} Gain \textbf{Ambassador Status} (convert a faith-based Position penalty into a bonus once per session).
    \item \textbf{The Ninth Bell Silent.} The village agrees to ring only eight bells, leaving the ninth for the old patrons. \textit{Effect:} Shrine-devout characters no longer suffer $-1$ Position in this village.
    \item \textbf{The Skald's Compromise.} A Temple hymn is set to an old shrine tune. \textit{Effect:} Any Audience tag that includes ``Suspicious'' downgrades to ``Curious'' in this region.
    \item \textbf{Blood-Silver to the Poor.} A feud is settled by paying wergild to the Temple's poor fund -- but the shrine witnesses the payment. \textit{Effect:} The settlement counts as both a \textbf{Thing Verdict} and a \textbf{Synod Absolution}. The GM may not use this feud as a complication again.
    \item \textbf{The Unbroken Spiral.} A new rite is born: a white candle and a raven's feather are offered together. \textit{Effect:} The crew gains a permanent \textbf{Synthesis Token} that refreshes every season (instead of being one-use).
\end{enumerate}

% Index entries (kept and expanded)
\index{Linn|see{Skerries and Storm-Oaths}}
\index{Theme generator}
\index{Skerries and Storm-Oaths generator}
\index{Places!Linn}
\index{People!Linn}
\index{Complications!Linn}
\index{Rewards!Linn}
\index{Sea-Queen of the Linn}
\index{High Jarl}
\index{Volva of the Mist}
\index{Thing-holm}
\index{Mistlands}
\index{Whale-road}
\index{Dolmis waystation}
\index{Runestone causey}
\index{Kelp-skerry gut}
\index{Aberderrin current}
\index{Thing-speaker}
\index{Skald}
\index{Mist-pilot}
\index{Net-wife}
\index{Oar-master}
\index{Shield-band}
\index{Foster-son}
\index{Storm-oath}
\index{Feud token}
\index{Raid-truce ribbon}
\index{Saint Adar}
\index{Adar the All-Father}
\index{Rorg the Storm Father}
\index{Saint Hervor}
\index{Tide-Washed Covenant}
\index{Draugr Jarl}
\index{Temple of Light!Linn}
\index{Frost Mark}
\index{Season's Turn}
\index{Wyrd}
\index{Tv\'{\i}h\"{o}gg}
\index{Volva}
