\section{Midh Ahkaz --- ``The Violet Gate''}
\label{chap:midh-ahkaz}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{Coin talks. Grass walks. Blood settles the rest.}

\begin{quote}
  \textbf{Khatun of the Ring (Ykrul Elder):}  
  ``The wind carries names, not tallies. But here, in the shadow of the Oasis Compact, both trade and oaths change hands. Do not mistake our patience for weakness. The grass remembers every hoof that crossed it.''
\end{quote}

\begin{quote}
  \textbf{Vilikari Factor (Night Market Regular):}  
  ``Ecktoria calls it uncivilized. We call it unbeatable. No writs, no seals, no questions. Coin talks. The rest walks. And if a snake offers you a sip from its cup, walk faster.''
\end{quote}

\noindent\textbf{Genre / Mood:} Frontier grit, steppe noir, law of the rope and the blade, sword and sorcery.

\noindent\textbf{Starting Location:} The Violet Bazaar at dusk --- torches gutter, braids slap against leather vests, and a one-eyed horse trader spits into the dust at your feet. ``You're new. The Oasis Well runs low, the Khatun's coughing blood, and someone's been passing hollow silver. Pick a problem --- or I'll pick one for you. And keep your hand off your steel near the Serpent Gate. The Coiled Ones have eyes everywhere.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=green!60!black,title={Theme \& Atmosphere}]
Midh Ahkaz is not a city. It is a wound that learned to trade. A day's ride north of the Ecktorian border, it squats on the edge of the Violet Steppe like a patient wolf. No walls. No gates. Just rings of felt tents, latrine trenches, and the smell of horse sweat, cedar smoke, and the faint copper tang of open claims.

The grass grows tall enough to hide a body. The locals cut it back at dawn --- but never at dusk, because the things that crawl through it at night have sharp teeth and no patience for property lines. Ecktorian emissaries call this place ``the Brink.'' Caravan guards call it home. Thieves call it paradise.

Every morning, the Bronze Court opens its scales under a canvas dome. Every night, the Dust Market auctions goods that fell off a wagon --- or off a corpse. The Khatun's word is law until someone stronger speaks. And in the Silver Stall Stable, a horse thief can buy a fresh mount with a story and a knife.

But a newer power has been coiling through the city's underbelly: the \textbf{Serpent Cult of Isoka}. They do not worship coin or grass. They worship \emph{shedding} --- the promise that what is weak can be burned away and reborn. Their initiates wear a single snake scale on a leather cord and speak in whispers. Their priests carry vials of venom that they call ``medicine.'' Their temple is a sunken pit behind the Dust Market, and at midnight the faithful drink from a cup that kills --- or transforms. The Khatun tolerates them because they offer their share of water. The Ykrul elders watch them with narrowed eyes. The snakes, they say, do not forget a face.

\vspace{2mm}
\textbf{What you notice first:} The constant low drumming of hooves --- a heartbeat you feel in your teeth. The way every merchant has a scale on their table \textit{and} a knife within reach. The braids in every horseman's mane, each knot a story, an oath, a threat. And the smell of something sweet and rotten near the Serpent Gate --- the cult's incense, or venom, or both.

\textbf{The one rule that kills newcomers:} Never refuse the guest-cup. The first pour is courtesy. The second is negotiation. The third is a reckoning unspoken. And the fourth --- the fourth means you're already dead, you just haven't stopped breathing. If the cup is offered by a gray-robed figure with a serpent tattoo on their palm, refuse it anyway. Run.
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!60!black,title={Western Steppe vs. Eastern Steppe (GM Sidebar)}]
The Ykrul divide their world into two great grass seas:

\medskip
\noindent\textbf{Western Steppe (Violet Prairie):}  
Rolling, fertile, forgiving. The grass grows tall enough to hide a crouched rider, but the land is laced with caravanserais, seasonal wells, and the smoke of permanent camps. This is where Midh Ahkaz sits --- the soft underbelly of the steppe, where trade happens and clans winter. The violet seed-heads give the prairie its name.

\medskip
\noindent\textbf{Eastern Steppe (True Steppe):}  
Vast, dry, harsh. Shared with the Kuvani horse-clans, it stretches east of the Valewood. Grass is shorter, winds are colder, and a rider can see a day's ride in every direction. The Ostrikari who appear in Midh Ahkaz are from the eastern fringes of the \textit{Western} Steppe --- they speak with a harder accent, braid their horses' tails differently, and carry bows of horn and sinew, not wood. Ykrul of the Eastern Steppe are rarely seen in the Western.

\medskip
\noindent\textit{Why this matters:} A Ykrul from the Eastern Steppe might find Midh Ahkaz almost claustrophobic. A trader from the Violet Prairie might feel lost in the east. Use the contrast to create tension between clans or to flavor foreign visitors.
\end{tcolorbox}

\subsection*{Regional Mood: Wind, Grass, and a Knife in the Back}

Midh Ahkaz thrives on movement --- and on the friction that movement creates. Caravans arrive from every direction: Ykrul horse-tribes from the deep steppe, Ostrikari mercenaries with horn bows and longer grudges, Vilikari traders who keep two sets of weights (one for the tax man, one for the grave), and the occasional bold Ecktorian factor with a silver tongue and a hidden dagger. The city's ``law'' is the Oasis Compact --- a shifting web of seasonal rights, grazing permits, and arbitration courts held under open sky. Disputes are settled by the Bowl (fairness) then the Board (geometry). A broken promise can cost a herd. A kept oath can buy a ford for a generation.

But when the Bowl spills and the Board cracks, the real law is the one you carry on your belt. The grass will hide your body --- or someone else's. And in the shadows, the Serpent Cult watches, waiting for the strong to devour the weak.

\noindent\textbf{Starting Location:} The Violet Bazaar, a sprawling market of felt tents and chalk circles, where horse traders shout over the jingle of brass bridles. A Vilikari merchant with oiled braids hands you a sealed letter. ``The Bronze Court suspects a false seal. Find the maker before the Grass Moon, or your shadow is marked for all to see. And if a woman in a snake-scale cloak offers you directions, take the long way.''

\paragraph*{(Ring/Well/Cairn)} The Oasis Well where water shares are drawn --- the token-keeper has a scarred face and a heavier hand than he should. Counting pebbles placed on stone ledges by memory, not sight. The distant smoke of a Steppe Ranger campfire --- or maybe a grass fire. Either way, it's moving toward you.

\section*{Spades --- Places}
\label{sec:midh-ahkaz-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Violet Bazaar} --- Felt stalls, chalk circles, horse traders and rumor merchants.
  \hfill \textit{The chalk circles are for disputes. Step inside and speak truth --- or settle in grass. Blood washes off the chalk, but the stain stays.} `[TRADE]`
\item \textbf{Bronze Court} --- Tariff court under a canvas dome; scales of bronze and iron.
  \hfill \textit{The scales were blessed by a dying Sky-Speaker. They tip even when the wind is still. The clerk has never smiled.} `[JUSTICE]`
\item \textbf{Oasis Well} --- Stone-rimmed pool, token-keeper with a leather bag of water rights.
  \hfill \textit{Each token is a carved pebble. Lose it, and your herd drinks dust. The keeper's knife is sharper than his tongue.} `[SURVIVAL]`
\item \textbf{Steppe Ranger Camp} --- Tent line on the eastern edge; trackers, scouts, feud mediators.
  \hfill \textit{The rangers wear wolf tails. They do not smile --- they wait. Their campfire is never allowed to go out.} `[WILD]`
\item \textbf{Caravan Rings} --- Concentric circles of wagon shelters; fires lit inward, watch posts on ox carts.
  \hfill \textit{The inner ring is for sworn allies. Outsiders sleep in the second ring, uneasy --- and a knife in the boot.} `[TRAVEL]`
\item \textbf{Sky-Speaker's Cairn} --- Stone heap atop a low hill; wind-knot ropes and bones.
  \hfill \textit{The shaman leaves offerings of horsehair. The wind carries his prayers to Kuva --- and sometimes his curses.} `[RITUAL]`
\item \textbf{Dust Market} --- No-questions auction at the city's edge; stolen goods, cursed relics, bad memories.
  \hfill \textit{The auctioneer wears a hooded mask. No one knows his face. No one has asked twice. The last man who did now sells his own teeth.} `[UNDERWORLD]`
\item \textbf{Silver Stall Stable} --- Remount station; horses for hire, gossip for free, blood for more.
  \hfill \textit{The stable master has three fingers on his left hand. He lost the others in a bet over a mare. He won the bet.} `[TRAVEL]`
\item \textbf{Grass Ring} --- Ritual dueling ground; mourners and bookies share benches.
  \hfill \textit{The grass grows back faster after a duel. The dead fertilize the peace. The watchers speak their peace before the body cools.} `[COMBAT]`
\item[J] \textbf{Foreign Quarter} --- Ecktorian, Vhasian, and Viterran merchants in permanent felt mansions.
  \hfill \textit{The Ecktorian envoy winters here. He has not left in five years. He has a name on a list in three countries --- but Midh Ahkaz doesn't extradite.} `[POLITICS]`
\item[Q] \textbf{Khatun's Ring} --- Largest tent, carpeted with woven battle scenes; the old chieftain's seat.
  \hfill \textit{The carpet is made of enemy cloaks. Each sash is a surrendered banner. The air smells of smoke, old blood, and the fear of petitioners.} `[AUTHORITY]`
\item[K] \textbf{Under-Arcades} --- Earthen tunnels below the bazaar; root-walled, bone-dusted.
  \hfill \textit{The walls are packed with bone dust and old burrows. The dead whisper route tips --- for a price. The trade is never silver.} `[STEALTH]`
\item[A] \textbf{Iron Well} --- Ancient Aeler shaft, now used for vengeance contracts and whispered names.
  \hfill \textit{The water tastes of rust and memory. Those who drink never lie again --- or never speak. The keeper writes the names of the condemned in ash. The ash floats.} `[DEATH]`
\item[\textbf{S}] \textbf{Serpent Pit (Cult Temple)} --- A sunken amphitheater behind the Dust Market, reeking of incense and old venom. `[MYSTERY]` `[DANGER]`
  \hfill \textit{The walls are lined with shed skins --- some as long as a horse. At midnight, the faithful drink from a bronze cup. Some rise renewed. Some do not rise at all. The cult's high priestess, a woman called Shedder, has not aged in twenty years.}
\end{enumerate}

\paragraph*{(Horse/Tally/Token)} Sky-Speaker's apprentice with a drum and a dream; Ostrikari arms-master selling short bows (he names each after a dead rival); Ykrul elder counting braids (her hair is a map of oaths).

\section*{Hearts --- People \& Factions}
\label{sec:midh-ahkaz-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Horse-Braid Apprentice} --- Young Ykrul with a head full of trade routes and a knot diary.
  \hfill \textit{She ties a new braid for every paying passenger. Her hair is a map of promises --- and grudges.} `[TRADE]`
\item \textbf{Ostrikari Arms-Master} --- Selling short bows and longer grudges; fights for coin.
  \hfill \textit{He names each bow after a dead rival. The bows never miss. His left ear is missing --- a score he settled in flesh.} `[COMBAT]`
\item \textbf{Bronze Court Clerk} --- Ink-stained, patient, scales in her eyes; never lies, never forgives.
  \hfill \textit{She records weights in blood. A single drop seals a verdict. She has never been wrong --- or so she says.} `[JUSTICE]`
\item \textbf{Vilikari Factor} --- Braided, oiled, quick with a smile and a second set of weights.
  \hfill \textit{His caravan has three names. Only the real one buys safe passage. The other two are for the tax man and the grave.} `[TRADE]`
\item \textbf{Steppe Ranger} --- Tracker with a wolf-tail cloak and a blood-oath to the Sky-Speaker.
  \hfill \textit{He never draws his bow unless watched. Hunger is his real weapon. His arrows are fletched with crow feathers.} `[WILD]`
\item \textbf{Tokan (Ykrul Scout)} --- Mute, sharp-eyed, always smelling the wind; her ears miss nothing.
  \hfill \textit{She signs with her knuckles. Three taps is ``danger.'' One tap is ``run.'' No taps means you're already dead.} `[STEALTH]`
\item \textbf{Ecktoria Emissary} --- Proud, wax-sealed, fuming at the ``primitive'' customs.
  \hfill \textit{He carries a copy of Imperial Code that he has never opened. The flies love his ink. The locals love his purse.} `[POLITICS]`
\item \textbf{Sky-Speaker's Shadow} --- Elder's quiet apprentice; listens more than the wind.
  \hfill \textit{She can name every clan's sky-omen. She cannot name her own origin. Her eyes are the color of ash.} `[RITUAL]`
\item \textbf{Knot-Keeper (Tulkani)} --- Counts caravan promises with a rope of nine strands.
  \hfill \textit{Each knot is a promise. Untie one and the bond dissolves. Cut one and the one who swore dies in his sleep. No one knows how.} `[SOCIAL]`
\item[J] \textbf{Dust Market Auctioneer} --- Hooded, genderless voice; knows the weight of every sin.
  \hfill \textit{The hammer is a thigh bone. The block is a skull altar from the Direwood. The skull's jaw moves when the price is right.} `[UNDERWORLD]`
\item[Q] \textbf{Khatun of the Ring} --- Gray-haired, one-eyed, still deadly at sixty; speaks for seven clans.
  \hfill \textit{She lost her eye to a Linn raider. She killed him with his own hook. She keeps the hook on her belt.} `[AUTHORITY]`
\item[K] \textbf{Kuva's Voice (Wandering Priest)} --- Weathered, ash-marked, offers omen readings for grass.
  \hfill \textit{He prays to the sky by throwing bones. The bones never lie --- they just mislead. The ash on his forehead is from a fire that hasn't been lit yet.} `[RITUAL]`
\item[A] \textbf{Iron Well Keeper (Vengeance Broker)} --- Mute, scarred, accepts names written in ash.
  \hfill \textit{A drink from the well costs a memory. A death costs a fortune. His mark is carved on his own arm --- the names move.} `[DEATH]`
\item[\textbf{S}] \textbf{The Shedder (Serpent Cult High Priestess)} --- A woman with snake-scale tattoos covering her arms, her voice a dry rustle.
  \hfill \textit{She offers the Cup of Renewal. Drink, and she promises you will shed your weakness. Some wake stronger. Some wake as snakes. She never drinks herself.} `[MYSTIC]` `[DANGER]`
\end{enumerate}

\paragraph*{(Drought/Feud/Blight)} Dry well cuts water rations --- the token-keeper has a list, and the list has a knife mark beside your name; blood-claim spills into the bazaar --- a boy's corpse lies in a chalk circle, and no one will claim it; locust swarm from the east --- the grass blackens, and the swarm leaves a sticky residue that smells of burnt honey and dread.

\section*{Clubs --- Complications/Threats}
\label{sec:midh-ahkaz-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Dry Well Ration} --- Water shares halved; prices spike, tempers fray. `[SURVIVAL]` `[SOCIAL]`
  \hfill \textit{The token-keeper has a list. The list has a knife mark beside your name. He didn't put it there.}
\item \textbf{Blood-Claim Feud} --- A killing in the Horse Ring; kinsmen demand redress in silver or flesh. `[WEIGHT]` `[COMBAT]`
  \hfill \textit{The victim's tribe drums the death rhythm. Every beat is a warning. The crowd is already wagering on the retaliation.}
\item \textbf{Locust Swarm} --- Grass blackens; caravans stranded, fodder worthless. `[WEATHER]` `[TRAVEL]`
  \hfill \textit{The swarm leaves a sticky residue. It smells of burnt honey and dread. The locusts spell words as they fly --- sometimes a name, sometimes a warning.}
\item \textbf{Hollow Silver} --- Light silver floods the Dust Market; Oasis Compact inquiry. `[WEIGHT]` `[JUSTICE]`
  \hfill \textit{The hollow coins are warm. The real ones are cold. Touch quickly. The maker has a scar on his left thumb.}
\item \textbf{Grass Fire} --- Wind-whipped flame from the east; evacuations, scorched trade paths. `[WEATHER]` `[DISASTER]`
  \hfill \textit{The smoke spells a name. Yours. Or your enemy's. Hard to tell. The fire moves against the wind.}
\item \textbf{Horse Sickness} --- Contagion at the Silver Stall Stable; quarantines, lost mounts. `[DEATH]` `[TRAVEL]`
  \hfill \textit{The sick horses dream. Their screams sound like your last argument. The stable master is weeping. He never weeps.}
\item \textbf{Khatun's Illness} --- Old chieftain fading; succession knives out. `[POLITICS]` `[WEIGHT]`
  \hfill \textit{Her death will split seven clans. The Sword Oath lasts one year. The poisoner is already savoring his reward.}
\item \textbf{Census Raid} --- Ecktorian legion patrol ``just visiting''; tensions high, bribes higher. `[AUTHORITY]` `[SOCIAL]`
  \hfill \textit{The patrol leader wears a mask of imperial wax. Pull it off and see a dead man's face. The face is yours --- from a life you haven't lived yet.}
\item \textbf{Sky-Omen (Red Wind)} --- Dust storm from the Ashaani desert; all navigation Desperate. `[WEATHER]` `[FEAR]`
  \hfill \textit{The dust tastes of iron and old blood. Breathe it and dream of falling. The wind whispers a name. Listen and you'll answer.}
\item[J] \textbf{Oath Challenge} --- A Kon'reh master demands geometry arbitration; your bond is the stake. `[WEIGHT]` `[SOCIAL]`
  \hfill \textit{The board is a painted skull. The pieces are horse teeth. The master hasn't lost in forty years. He's due.}
\item[Q] \textbf{Court Strike} --- Bronze Court clerks refuse to speak the measures; trade freezes, tempers explode. `[JUSTICE]` `[TRADE]`
  \hfill \textit{The clerks chant a measuring song. Stop them before they reach nine. The ninth note is not a sound. It is a name.}
\item[K] \textbf{Duel for Rights} --- Two clans claim the same well; witnesses needed to prove water share. `[COMBAT]` `[JUSTICE]`
  \hfill \textit{The duel is fought with blunt spears. First blood wins the well. Both bleed. The well doesn't care.}
\item[A] \textbf{Kuva's Vengeance} --- The sky itself turns; hail, lightning, and a voice from the clouds. `[DEATH]` `[PATRON]`
  \hfill \textit{The voice names a sinner. Your name. The hail has teeth. The lightning writes your confession in the grass.}
\item[\textbf{S}] \textbf{Serpent's Due} --- Cult enforcers demand a ``gift'' for the Shedder. Refuse, and you may wake with a viper in your saddlebag --- or in your bed. `[DANGER]` `[MYSTERY]`
  \hfill \textit{They do not ask for coin. They ask for a name, a secret, or a hair from a purebred horse. The Shedder weaves these into her nest.}
\end{enumerate}

\paragraph*{(Brand/Token/Pass)} Grass Right for one season of grazing --- a branded stick that warms when the grass is healthy; Oath-Bead to compel a Kon'reh master's arbitration --- the bead is a single horsehair knot, tightening when the master speaks truth; Serpent Scale (cult favor) --- a single iridescent scale that can buy one question from the Shedder, but marks you as her prey.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:midh-ahkaz-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Grass Right} --- One season of grazing on a named meadow; herders pay you tribute. `[TRAVEL]` `[SOCIAL]`
  \hfill \textit{The right is a branded stick. Burn it to end the lease --- or renew. The brand is a wolf's head. It howls when the meadow is overgrazed.}
\item \textbf{Oath-Bead} --- A single knot of horsehair; compels a Kon'reh master to arbitrate your dispute. `[JUSTICE]`
  \hfill \textit{The bead grows warm when the master speaks truth. Cold when he lies. He has never felt it grow cold.}
\item \textbf{Water Share} --- Priority access to a well during drought; token from the Oasis Compact. `[SURVIVAL]`
  \hfill \textit{The token is a polished pebble. Drop it in the well to claim your draw. The well will remember your face.}
\item \textbf{Steppe Ranger Favor} --- One dangerous tracking job resolved off-screen. `[WILD]`
  \hfill \textit{The ranger marks your name on his tent pole. He never forgets a favor. He also never forgets a slight.}
\item \textbf{Vilikari Voucher} --- No-questions cargo space on three caravans; smuggle as you will. `[TRADE]` `[STEALTH]`
  \hfill \textit{The voucher is a strip of oiled leather. The factor's scent guides the mules. The mules are smarter than the factor.}
\item \textbf{Census Waiver} --- Ecktorian patrol leaves you alone for one crossing; paper shield. `[AUTHORITY]`
  \hfill \textit{The waiver is stamped with a bear. The bear's eyes follow you. The stamp is a lie. The waiver still works.}
\item \textbf{Khatun's Whistle} --- Summon a small warband once (Cap 3) for a single scene. `[COMBAT]`
  \hfill \textit{The whistle is a hollow horse tooth. Blow it and the grass shivers. The riders come. They expect to be repaid in stories.}
\item \textbf{Sky-Speaker's Omen} --- One weather forecast that cannot fail; plan your escape. `[WEATHER]`
  \hfill \textit{The omen is a bone thrown into fire. Read the cracks --- carefully. The fire will answer. It will also ask.}
\item \textbf{Silver Stall Chit} --- One free remount exchange; never be horseless. `[TRAVEL]`
  \hfill \textit{The chit is a brass hoof. Present it and the stable master nods once. The horse will be fast. The horse may also be cursed.}
\item[J] \textbf{Burned Mark} --- A written promise from a rival, witnessed by three camp elders. `[WEIGHT]` `[SOCIAL]`
  \hfill \textit{The mark is burned into a leather strip. Wash it and the vow dissolves. Burn it and the rival dies in his sleep. No one knows why.}
\item[Q] \textbf{Dust Market Share} --- A portion of the auctioneer's commission for a season; wealth in sin. `[TRADE]`
  \hfill \textit{The share is paid in mismatched coins. Spend them fast before they curse you. The auctioneer's mask has a new scratch every week.}
\item[K] \textbf{Kin-Bound Bond} --- A criminal clan owes you a life; they will kill once for you --- or die trying. `[DEATH]` `[WEIGHT]`
  \hfill \textit{The bond is a braid of enemy hair. Cut it to call the blood. The clan's matriarch will come. She will ask whose.}
\item[A] \textbf{Seal of the Oasis Compact} --- Casting vote on any water right dispute. Once. `[JUSTICE]` `[AUTHORITY]`
  \hfill \textit{The seal is a bronze well with a snake curled at the bottom. Turn it and the water booms. The snake will bite if you turn it back.}
\item[\textbf{S}] \textbf{Cup of Renewal (Cult Favor)} --- A single sip from the Shedder's cup. You may reroll any failed Body or Spirit roll once, but gain a permanent snake-scale mark on your palm. The mark will itch when the cult calls. `[MAGIC]` `[DANGER]`
\end{enumerate}

\section*{Quick use notes}
\label{sec:midh-ahkaz-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are narrative permits (grazing rights, oaths, witnesses) that shift Position rather than demand a roll.
\item If any A appears, echo \textbf{grass and wind} motifs --- shifting alliances, old claims that ride with the seasons, the sky as witness, a body that vanishes into the tall grass by morning.
\item The violet seed-heads of the steppe are a recurring omen: when they glow under moonlight, a spirit walks; when they shed, a caravan will be late --- or never arrive.
\item If a \textbf{Serpent} (S) entry appears, introduce a cult encounter, an omen of shedding, or a venomous complication.
\end{itemize}

\subsection*{Additional Features (Frontier Grit)}
\label{sec:midh-ahkaz-mechanics}

\begin{itemize}
\item \textbf{Kon'reh Arbitration:} Disputes are settled on a board painted on hide. Winners gain seasonal rights. Losers yield grass or labor --- or blood if they have neither left to give. The board is set up in the Grass Ring. The spectators bet on the outcome.
\item \textbf{Grass Omens:} The steppe flickers violet when rain is near. When the seed-heads droop at noon, a blood feud approaches. When they stand straight at midnight, the Hollow walks. No one stays outside when the seed-heads stand straight.
\item \textbf{Khatun's Braid:} The chieftain's hair holds the memory of every oath she swore. Cutting a braid voids a promise --- but the severed braid becomes a curse token. The current Khatun has nine braids. She is saving the tenth for something important.
\item \textbf{The Law of the Rope:} Midh Ahkaz has no jail. Serious offenders are tied to a post in the Grass Ring and left for the steppe. If they survive the night, they are free --- and must answer to the family of their victim. No one has survived the night in seven years.
\item \textbf{The Shedding (Cult Rite):} Once per month, the Serpent Cult holds the Rite of the Cup. Participants who survive the venom gain +1 die to resist poison and disease for one month, but also gain a random quirk (nightmares, a phobia of open flames, an insatiable hunger for raw meat). The GM may offer this as a bargain --- but the cult always asks for something in return later.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Wind Father} --- The Traveler. \textit{He who reads the grass and knows the safe routes --- for a share of the cargo.}
    \item \textbf{The Scale Mother} --- Mykkiel. \textit{She who weighs the oath and the old score; her clerks carry knives.}
    \item \textbf{The Grass King} --- Grimmir. \textit{The wild that pushes back against the ring of tents; he demands a sacrifice of horseflesh each spring.}
    \item \textbf{The Toothless One} --- Malachai. \textit{A silver coin in the dust market that guarantees a fair deal --- until the grass remembers.}
    \item \textbf{The Shedder (Cult)} --- Isoka. \textit{She who teaches that weakness can be burned away. Her cup is poison. Her mercy is a fang.}
    \item \textbf{Crom (Vilikari Barbarian Cult)} --- Unknown. \textit{A few Vilikari outcasts worship a god of uncivilized strength, called Crom. They say he sits on a mountain and does nothing, but that he laughs when you kill your enemy with your bare hands. His shrine is a pile of skulls behind a ruined bathhouse. No one knows if he exists. No one argues with his worshippers.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Bitter Root:} Daughters of the Cord run a ``free soup'' wagon near the Dust Market. They ask no coin --- but they measure your shadow. A renegade Sister hides among the horse-braid apprentices, trying to untie her cord from a blood feud.
  \item \textbf{The Threshold Compendium:} Saikou Ira once tracked a ghoul to the Under-Arcades. He left a silver mirror in the wall. It still shows the ghoul's face. The ghoul is still watching.
  \item \textbf{Malachai, the Chained Angel:} A broken Kon'reh board at the Iron Well is said to compel truth. Players who touch it gain +1 die to resist lies --- but mark 1 Obligation to Malachai. The board was broken by a merchant who lied about a horse. The horse was a demon.
  \item \textbf{The Grey Wanderer's Grimoire:} Midh Ahkaz psions are called ``Wind-Readers.'' They sense lies by the tremor in grass. The Khatun employs two; they never speak to each other. They communicate by writing in the dust. The dust blows away after each message.
  \item \textbf{Isoka's Serpent Cult:} The cult's venom is said to be harvested from a great snake that lives in a sealed cave beneath the Iron Well. The snake is called the First Shedder. It has not eaten in three hundred years. It does not need to. It waits.
  \item \textbf{The Ninth:} In any Midh Ahkaz scene, the ninth horse table in the Violet Bazaar sells nothing but bad advice. Spend a Boon to reroll a failed negotiation --- but the advice may haunt you. The table is always empty except at midnight. At midnight, a hooded figure sits there. No one knows who.
\end{itemize}

\subsection*{Midh Ahkaz Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Hoofprint Pour:} Pour a cup of water into the first hoofprint of a new horse. Ignore the first \emph{Horse Sickness} penalty. `[SUPERSTITION]`
  \item \textbf{Braid the Knot:} Tie a strand of mane hair around your wrist. Cancel \emph{Counterfeit Coin} or take --1 die on your next Charisma roll (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Grass:} Whisper an oath-breaker's name into the wind at noon. Gain +1 Effect when collecting from them this scene, but mark \emph{Obligation} to a Sky-Speaker. `[SUPERSTITION]`
  \item \textbf{Scale on the Tongue (Cult superstition):} If a serpent crosses your path, spit three times. The next venomous bite against you automatically misses --- but the Shedder will know your name. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table (Midh Ahkaz)}
\label{sec:midh-ahkaz-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw attention from \textbf{Steppe Rangers or Bronze Court clerks}.
\item \textbf{Noise:} The sound carries across the grass; a \textbf{rival clan's scout} overhears.
\item \textbf{Trace:} Hoofprints, broken grass, or losing an arrow marks your path.
\item \textbf{Delay:} A water ration takes longer; a horse tires without rest.
\item \textbf{Grass Strain:} Mark +1 segment on a relevant \textbf{resource clock}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Border riders} converge on your position; a horn sounds.
\item \textbf{Position Lost:} You lose high ground/cover/the wind due to a \textbf{grass fire or horse stampede}.
\item \textbf{Foe Appears:} A \textbf{Vengeance broker, Ostrikari sell-sword, or clan champion} arrives.
\item \textbf{Gear Trouble:} A \textbf{bridle breaks, bow string snaps, waterskin leaks}.
\item \textbf{Oath Hearsay:} A false rumor about your oath reaches the wrong ear.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{horsemen, kin-claim hunters, or Dust Market thugs} arrive.
\item \textbf{Key Gear Breaks:} A \textbf{horse goes lame, seal ring cracks, bound-mark burns}.
\item \textbf{Major Twist:} The situation changes --- \textbf{Khatun dies, well runs dry, duel interrupted by grass fire}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} (e.g., Blinded by dust, Parched).
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A \textbf{hidden pit, poisoned water share, or false oath} activates.
\item \textbf{Authority Arrival:} \textbf{Khatun's warband, Ecktorian patrol, Sky-Speaker, or the Shedder} intervenes.
\item \textbf{Scene Shift:} The environment changes --- \textbf{grass fire sweeps through, dust storm blinds, night falls without warning}.
\item \textbf{Patron Omen:} A \textbf{bolide, strange animal, or wind that speaks your name} appears.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn --- \textbf{the real falsifier was your ally; the water right was a trap.}
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Midh Ahkaz (Grass \& Trade):} Grass catches fire, horses spook at shadows, bargains settled with a knife on the table, violet seed-heads floating in the wind.
\item \textbf{Midh Ahkaz (Oath \& Sky):} Clouds spell forgotten names, wind carries false testimony, a duel's outcome shifts with the sun, a violet dusk signals a spirit.
\item \textbf{Midh Ahkaz (Horse \& Bow):} An arrow splits in flight, a horse refuses a rider, a bowstring sings a dying clansman's tune, a braid comes undone at the wrong moment.
\item \textbf{Midh Ahkaz (Serpent Cult):} A shed snakeskin is found in your tent; a cup of warm milk appears on your saddle; a whisper from the dark says, ``The Shedder invites you to dine.''
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Ride):} On a hit, shift one band. On a strong hit, also impose \emph{Dust Cloud}.
\item \textbf{Cut the Herd (Presence + Sway):} On a hit, freeze range for one exchange; if a Grass Right is in your possession, +1 die.
\item \textbf{Weather the Stampede (Resolve + Survival):} On a hit in \emph{Horse Sickness}, you pick which mount survives; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (three-phase foes)}
\begin{itemize}
\item \textbf{The Hollow Scale (cursed scale bound to the Oasis)} --- \emph{Phases}: Fair Measure $\rightarrow$ Thirst Unearned $\rightarrow$ Naming the Bound. \emph{Cracks}: burn a false token in the Oasis well; satisfy a seven-clan water claim; let a Sky-Speaker rebless the scale.
\item \textbf{Grass Reaver (vengeful horse-lord, half-legend)} --- \emph{Phases}: Hoof Drum $\rightarrow$ Ghost Herd $\rightarrow$ Blood-Claim Ride. \emph{Cracks}: confess a broken oath at his cairn; return a stolen foal; outrun him on a borrowed horse.
\item \textbf{The Hollow Merchant (greedy soul who sold his reflection)} --- \emph{Phases}: Roll of the Lost $\rightarrow$ Chain of Measures $\rightarrow$ Break the Seal. \emph{Cracks}: settle a forgotten claim; break his bond before witnesses; let the Iron Well drink his name.
\item \textbf{The Shedder (Serpent Cult Ascendant)} --- \emph{Phases}: Invitation $\rightarrow$ The Cup $\rightarrow$ Transformation. \emph{Cracks}: refuse the cup three times; show her a scale from the First Shedder; cut her braid --- she will shed her skin and flee, but she will remember your face.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power in Midh Ahkaz}]
In Midh Ahkaz, power flows through horseflesh, water rights, and the geometry of oaths. The Khatun rules by consensus, but the true authority rests with the Oasis Compact --- a loose council of clan elders, Vilikari factors, and Steppe Rangers who interpret the Wind Law. When the Compact fails, the Grass Ring becomes the courtroom, and the duel is the verdict.

But a new power coils beneath the old: the Serpent Cult of Isoka. They offer a different kind of power --- transformation through poison, renewal through death. The weak are shed like skin; the strong rise reborn. The cult does not seek to rule Midh Ahkaz. It seeks to own the people who will rule after the old ways die.

\textbf{For the GM --- Actionable Examples:}
\begin{itemize}
\item A player who wins a Kon'reh match earns a \textbf{Grass Right} (Diamond). That right could be contested by a rival clan the very next session --- forcing a rematch or a chase across the prairie. The grass will be cut for the rematch. The loser's grass will be salted.
\item A \textbf{Water Share} token is stolen. The thief leaves a violet seed-head as a signature --- and a single snake scale. Tracking them leads to a seasonal camp where the Serpent Cult is holding a secret Rite of the Cup. The party must decide whether to interfere --- or to drink.
\item The Khatun's illness (Club) is not natural. A Vilikari factor (Heart) knows the poisoner's name but demands a \textbf{Burned Mark} (J$\diamondsuit$ Diamond) in exchange --- a dangerous witness the player must then protect. The witness is a child who cannot speak. She draws in the dust. Her drawings keep showing a serpent eating its tail.
\end{itemize}
These hooks turn the generator entries into immediate playable situations.
\end{tcolorbox}

% Index entries
\index{Midh Ahkaz|see{Violet Gate}}
\index{Theme generator}
\index{Steppe generator}
\index{Places!Midh Ahkaz}
\index{People!Midh Ahkaz}
\index{Complications!Midh Ahkaz}
\index{Rewards!Midh Ahkaz}
\index{Khatun of the Ring}
\index{Violet Bazaar}
\index{Bronze Court}
\index{Oasis Well}
\index{Steppe Ranger}
\index{Grass Ring}
\index{Kon'reh arbitration}
\index{Sky-Speaker}
\index{Iron Well}
\index{Tokan}
\index{Serpent Cult of Isoka}
\index{The Shedder}
\index{Crom (Vilikari deity)}