\section{Mistlands --- ``Bells, Salt, and Breath''}
\label{chap:mistlands}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Mistlands lie north of the Aelerian mountains, where the Aberderrin Sea laps against reed-fens and drowned shores. To the west, the Direwood breathes; to the east, the salt tides of the Aberderrin. It is a Protectorate of the Aeler, but their iron passes only hold the southern roads --- the land itself belongs to fog and bells. Every levee hums with ward-notes, every ferry takes salt as tithe, and every shadow is measured against the Direwood's endless breath. The people here live by the weight of salt, the echo of bells, and the hope that their names will not be spoken by the mists.

The Aeler Protectorate maintains order with patrols, garrisons, and breath-bonds, but the real law is older: salt lines drawn at dusk, iron nails hammered into threshold posts, and the unspoken knowledge that the ninth bell must never ring. The Mistlands do not forgive forgetfulness. The mist has a long memory, and it shares its grudges freely.

\vspace{2mm}
\textbf{What you notice first:} The constant, distant chime of bells --- never quite in sync. The taste of salt on your lips even inland. The way everyone holds their breath when crossing a bridge. The silence when a bell fails to ring.

\textbf{The one rule that kills newcomers:} Never answer a voice that calls your name from the fog. The mist collects names, and once it has yours, the ferryman will always know where to find you.
\end{tcolorbox}

\subsection*{Regional Mood: Bells and Breath}

Here, breath is currency and silence a sin. Every exhalation is counted by the ward-bells, and every stolen inhalation feeds something that should not be. The Aeler Protectorate maintains order with patrols and garrisons, but the real law is older: salt lines drawn at dusk, iron nails hammered into threshold posts, and the unspoken knowledge that the ninth bell must never ring. The Mistlands do not forgive forgetfulness. The mist has a long memory, and it shares its grudges freely.

The tension between the Aeler governor (Legate Halvar) and the local bell-wardens is constant. The Aeler want standardized bell-codes, the salt-tithe weighed true, and villages who keep the old pacts. The wardens want the old ways: counting in eights, never the ninth, and the right to refuse a name to the fog. Payden's Port, the only true city, is a battleground of writs and lanterns. Grain ships leave each autumn, carrying the Protectorate's due east and returning with iron, engineers, and the occasional missionary --- who rarely stay.

\subsection*{Prominent NPCs of the Mistlands}
\label{sec:mistlands-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Legate Halvar & Aeler governor of the Protectorate & Keep the bell-lines intact and the grain flowing east & She carries a bell that has never been rung. It will ring once, at the end of her reign. \\
\hline
Mother Anja & Salt-Monk of the White Pans & Preserve the old wards against the Direwood & Her tongue is white with salt; she speaks only in blessings or silence. \\
\hline
Warden Lucas & Bell-Warden of the High Line & Protect his village from the mist and the Aeler law-men & He refuses to write down the ninth note; it lives only in his fading memory. \\
\hline
Sister Darya & Healer of the Grey Hospice of the Unnamed (Temple of Light) & Tend the fog-sick and the drowned & Her grey robes never stain, and she asks no payment --- only a name to carry in her keeping. \\
\hline
Notary Voss & Covenant judge (Mykkiel) & Impose Aeler law on local customs & He carries a water-timer to measure the shifting of legal moods. The mist ignores it. \\
\hline
The Ferryman & Oath-ferryman of the Ghost Slip & Gather tolls in salt and whispered names & His left hand is always gloved; underneath, the names are carved into his palm. \\
\hline
\end{longtable}

\subsection*{Names of the Mistlands}
\label{sec:mistlands-names}

Inspired by the coastal fens and the Aeler protectorate: clipped, breathy, with soft vowels and hard endings. Surnames often denote a marsh, a craft, or a ward.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Halvar & Anja & Foss & \textit{the Bell-Sworn} \\
\hline
Lucas & Darya & Mork & \textit{Fog-Born} \\
\hline
Voss & Pia & Saltvedt & \textit{the Unnamed} \\
\hline
Torben & Lise & Kyst & \textit{Breath-Thief} \\
\hline
Stian & Ragna & Dyrep & \textit{Ward-Keeper} \\
\hline
Vidar & Signe & Tangen & \textit{Ash-Finger} \\
\hline
\end{longtable}

\paragraph*{(Fen/Levee/Bell-line)} Reed-fen causeway with chalked ward-runes on every milestone; Bell-Line levee with iron posts and eight bells per span.

\section*{Spades --- Places}
\label{sec:mistlands-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Reed-Fen Causeway} --- Chalked ward-runes on every milestone; safe path through the mist. \card{\spadesuit}{TRAVEL}
  \hfill \textit{Step only on the chalked stones. The mud between has no bottom.}
\item \textbf{Bell-Line Levee} --- Iron posts, eight bells per span; one bracket empty, one note forbidden. \card{\spadesuit}{WARD}
  \hfill \textit{The empty bracket is for the bell that never rings. The rope is still there.}
\item \textbf{Ghost-Ferry Slip} --- Rope crosses black water; toll taken in salt and whispered names. \card{\spadesuit}{DEATH} \card{\spadesuit}{TRAVEL}
  \hfill \textit{The ferryman never blinks. Some say he traded his eyelids for the right to collect.}
\item \textbf{Pall Watch-Tower} --- Piles above mire; mirrors face west to the Direwood's breath. \card{\spadesuit}{WARD}
  \hfill \textit{The mirrors are silvered on both sides. The wardens watch their own reflections as much as the trees.}
\item \textbf{Mist-Chapel} --- Stilted planks; votive lanterns gutter under damp rafters. \card{\spadesuit}{RITUAL}
  \hfill \textit{The icon has no face. Worshippers say it lost its features to the fog, one by one.}
\item \textbf{Dead-Cut Canal} --- Old imperial ditch; patrolled by eel boats and law. \card{\spadesuit}{TRAVEL}
  \hfill \textit{The eels have eyes like men. Fishermen throw back any that blink.}
\item \textbf{Drowned Copse} --- Root-tangles clutch tide like old hands; bones in the bark. \card{\spadesuit}{DEATH}
  \hfill \textit{The trees were planted on a mass grave. They grew twisted, hungry.}
\item \textbf{Fogmill Ridge} --- Wind vanes vanish into cloud; gears hiss like prayers to fog. \card{\spadesuit}{WEATHER}
  \hfill \textit{The miller grinds salt, not grain. The fog claims his breath each turning.}
\item \textbf{Salt-Pan Terraces} --- Patched with peat; ward-salt dries under nets, watched by monks. \card{\spadesuit}{RESOURCE}
  \hfill \textit{The monks chant without ceasing. If they stop, the salt spoils --- and so does the warding.}
\item[J] \textbf{Fort-Stair} --- Stacked timber and stone; flags swallowed by mist, guards by routine. \card{\spadesuit}{AUTHORITY}
  \hfill \textit{The flags are changed daily. The ones that vanish are never spoken of again.}
\item[Q] \textbf{Witchlight Bridge} --- Over black runnel; ninth plank ``counted'' but unseen. \card{\spadesuit}{THRESHOLD}
  \hfill \textit{The missing plank is there --- you just can't see it. Step where it should be, or don't step at all.}
\item[K] \textbf{High-Mist Pass} --- Through Aelerians; cairns ringed in iron nails, bells in the stones. \card{\spadesuit}{TRAVEL}
  \hfill \textit{Each cairn holds a bell that rang once at the death of a warden. They ring again when mist takes another.}
\item[A] \textbf{Weeping Gate} --- Western palisade facing Direwood; bells wired to bones of trees. \card{\spadesuit}{WARD} \card{\spadesuit}{DEATH}
  \hfill \textit{The gate weeps sap that smells of iron and old grief. Wardens collect it in silver cups.}
\end{enumerate}

\paragraph*{(Warden/Ferryman/Aeler)} Bell-warden whose ringbook decides which notes keep which ghosts; oath-ferryman gathering toll in salt and names.

\section*{Hearts --- People \& Factions}
\label{sec:mistlands-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Reed-Cutter} --- Peat-black hands, precise opinion on fog; knows every safe path. \card{\heartsuit}{TRAVEL}
  \hfill \textit{She cuts reeds in a pattern --- one snip, two snips, then a pause for the ninth.}
\item \textbf{Salt-Monk} --- Blesses pans, keeps ward-measure honest; salt as both shield and vow. \card{\heartsuit}{RITUAL}
  \hfill \textit{His tongue is white with salt. He speaks only in blessings or silence.}
\item \textbf{Bell-Warden} --- Ringbook decides which notes keep which ghosts at bay. \card{\heartsuit}{WARD}
  \hfill \textit{She refuses to write down the ninth note. It lives only in her memory --- and that memory is fading.}
\item \textbf{Oath-Ferryman} --- Toll taken in salt and names; remembers both, gathers both. \card{\heartsuit}{DEATH} \card{\heartsuit}{SOCIAL}
  \hfill \textit{His left hand is always gloved. Underneath, the names are carved into his palm.}
\item \textbf{Lantern Acolyte} --- Trims wicks and arguments; light as law in the mist. \card{\heartsuit}{AUTHORITY}
  \hfill \textit{He carries a bell on his belt. It rings once for each lie he hears.}
\item \textbf{Mist-Pilot} --- Linn refugee; ghosts ships through reed-lace --- if the cost is breath. \card{\heartsuit}{TRAVEL}
  \hfill \textit{She never speaks above a whisper. Her voice is currency, and she spends it carefully.}
\item \textbf{Protectorate Clerk} --- Aeler oilskin folios, narrow smile; law in triplicate. \card{\heartsuit}{AUTHORITY}
  \hfill \textit{His stamp is a bell. He presses it twice --- once for the crown, once for the Hollow.}
\item \textbf{Shroud-Diver} --- Harvests tokens from drowned fields before dawn's false light. \card{\heartsuit}{STEALTH} \card{\heartsuit}{DEATH}
  \hfill \textit{He wears a bell around his ankle. If it rings, he's stayed too long.}
\item \textbf{Direwood Refugee} --- Cold breath, colder stories; knows what the trees whisper. \card{\heartsuit}{FEAR}
  \hfill \textit{She carved out her own tongue. The trees still speak through her --- in a voice not her own.}
\item[J] \textbf{Fog-Knight} --- Cavalry rides levee crowns by bellcode; mist as their steed. \card{\heartsuit}{AUTHORITY} \card{\heartsuit}{COMBAT}
  \hfill \textit{Their horses have no shadows. Some say they traded them for fog-sight.}
\item[Q] \textbf{Legate of Mists} --- Aeler governor; equal parts mercy and decree, bells at their command. \card{\heartsuit}{AUTHORITY}
  \hfill \textit{She carries a bell that has never been rung. It will ring once, at the end of her reign.}
\item[K] \textbf{Lord Warden} --- Local strong hand sworn to Legate's seal; law walks the fens. \card{\heartsuit}{AUTHORITY}
  \hfill \textit{He lost an eye to a wight. He wears a bell in the socket.}
\item[A] \textbf{Mist-Seer} --- Bargains breath for answers; candles shorten when she speaks. \card{\heartsuit}{MAGIC} \card{\heartsuit}{FEAR}
  \hfill \textit{She has no shadow. The mist took it as a pledge.}
\end{enumerate}

\paragraph*{(Undead/Weather/Law)} Ground-mist lifts off fen and eats tracks by the yard; witchlights try to count you; step wrong and bridge forgets you.

\section*{Clubs --- Complications/Threats}
\label{sec:mistlands-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Ground-Mist} --- Lifts from fen; eats tracks by the yard, names by the breath. \card{\clubsuit}{WEATHER} \card{\clubsuit}{DEATH}
  \hfill \textit{Your footfalls vanish behind you. The mist is following, not rising.}
\item \textbf{Witchlight Count} --- Trying to count you; step wrong and bridge forgets your path. \card{\clubsuit}{MAGIC} \card{\clubsuit}{FEAR}
  \hfill \textit{The lights flicker in a pattern --- two short, two long. They're signaling someone.}
\item \textbf{Ward-Salt Short} --- Pans go green; bells ring, nothing answers from the other side. \card{\clubsuit}{RESOURCE} \card{\clubsuit}{WARD}
  \hfill \textit{The salt is grey, not white. Someone cut it with grave dust.}
\item \textbf{Wrong Bell} --- Cracked note opens a door the Legate can't close; mist takes the gap. \card{\clubsuit}{MAGIC}
  \hfill \textit{The bell's clapper is a finger bone. The original owner wants it back.}
\item \textbf{Linn Raid} --- Slips reed-maze; hearths douse, horns travel on the fog. \card{\clubsuit}{COMBAT}
  \hfill \textit{The raiders wear bells on their belts --- stolen from the last village.}
\item \textbf{Direwood Moan} --- At dusk; Weeping Gate ropes go taut by themselves. \card{\clubsuit}{DEATH} \card{\clubsuit}{FEAR}
  \hfill \textit{The moan has words. You don't want to understand them.}
\item \textbf{Valewood Wind} --- Crosses Aberderrin; voices sweet, landmarks false. \card{\clubsuit}{WEATHER} \card{\clubsuit}{MAGIC}
  \hfill \textit{The wind smells of honey and rot. Do not follow it.}
\item \textbf{Bell-Line Failure} --- Levee breaks silence; wraith steps across like it owns the road. \card{\clubsuit}{WARD} \card{\clubsuit}{DEATH}
  \hfill \textit{The wraith carries a lantern. The light is blue --- the color of drowned breath.}
\item \textbf{Marsh-Quake} --- Sod heaves, dyke slumps; causeways and bodies sink together. \card{\clubsuit}{DISASTER}
  \hfill \textit{The ground exhales. Your name rises from it like smoke.}
\item[J] \textbf{Protectorate Interdiction} --- All ferries sealed ``until review''; routes dry up. \card{\clubsuit}{AUTHORITY} \card{\clubsuit}{TRAVEL}
  \hfill \textit{The seals are stamped with a bell. The wax is mixed with bone ash.}
\item[Q] \textbf{Rite-Purge} --- Inspectors call local customs heresy; tempers at a boil. \card{\clubsuit}{AUTHORITY} \card{\clubsuit}{SOCIAL}
  \hfill \textit{The inspectors carry blank books. They fill them with names of the accused.}
\item[K] \textbf{General Alarm} --- Pall horns sound; every lantern becomes law, every shadow suspect. \card{\clubsuit}{AUTHORITY} \card{\clubsuit}{FEAR}
  \hfill \textit{The horns are made from the tusks of a sea beast. The beast's ghost is angry.}
\item[A] \textbf{Tide-Mist} --- Rolls from Aberderrin; spirits ride the white and won't be named. \card{\clubsuit}{WEATHER} \card{\clubsuit}{DEATH}
  \hfill \textit{The spirits have no faces --- only the faces of the drowned.}
\end{enumerate}

\paragraph*{(Pass/Seal/Token)} Ward-salt allotment draws safe circle once; ferry token crosses named black runnel regardless of curfew.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:mistlands-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Ward-Salt} --- Pure measure draws safe circle once; mist parts, briefly. \card{\diamondsuit}{WARD}
  \hfill \textit{The salt is warm. It knows it will be used.}
\item \textbf{Ferry Token} --- Cross named black runnel regardless of curfew or tide. \card{\diamondsuit}{TRAVEL}
  \hfill \textit{The token is a knuckle bone. The ferryman accepts it without comment.}
\item \textbf{Bell-Key} --- Unlock and set one bell on the Line to your note; change the song. \card{\diamondsuit}{MAGIC}
  \hfill \textit{The key is a tuning fork. Strike it once to claim your note.}
\item \textbf{Lantern Writ} --- Lawful travel at night along levee crown with escort. \card{\diamondsuit}{TRAVEL} \card{\diamondsuit}{AUTHORITY}
  \hfill \textit{The writ is written on dried bladder. It crinkles when you lie.}
\item \textbf{Exorcist's Seal} --- Stamped clause downgrades ``haunt'' to ``nuisance.'' \card{\diamondsuit}{MAGIC}
  \hfill \textit{The seal is a bell without a clapper. It's silent, like the wraiths it binds.}
\item \textbf{Codeleaf} --- Today's mirror and horn sequences; fog cannot hide what you know. \card{\diamondsuit}{TRAVEL}
  \hfill \textit{The leaf is soaked in brine. Read it before it dries.}
\item \textbf{Protectorate Mark} --- Priority rations and rope from Fort-Stair stores. \card{\diamondsuit}{RESOURCE}
  \hfill \textit{The mark is a brand. It smokes when you show it.}
\item \textbf{Wraith-Pardon} --- Pardon for desperate deeds done while spirits press near. \card{\diamondsuit}{AUTHORITY}
  \hfill \textit{The pardon is signed in salt water. It will dissolve if you misuse it.}
\item \textbf{Bone-Field License} --- Salvage rights at drowned hamlet (with witnesses). \card{\diamondsuit}{RESOURCE}
  \hfill \textit{The license is a map drawn on a shroud. Follow the stitch marks.}
\item[J] \textbf{Refuge-Right} --- Move a named household through Gate unchallenged; breath as token. \card{\diamondsuit}{SOCIAL}
  \hfill \textit{The right is a bell-rope. Tie it to your belt. Nonna will ask.}
\item[Q] \textbf{Legate's Audience} --- One whisper exits as policy; bells ring your decree. \card{\diamondsuit}{AUTHORITY}
  \hfill \textit{The Legate's ear is a silver bell. Speak into it once.}
\item[K] \textbf{Warden's Commission} --- Raise levee guards; local doors open, and trust follows. \card{\diamondsuit}{AUTHORITY}
  \hfill \textit{The commission is a whistle. Blow it, and the fens answer.}
\item[A] \textbf{Pall Indulgence} --- Temporary exception to ward law, spoken and sealed by bell. \card{\diamondsuit}{MAGIC} \card{\diamondsuit}{AUTHORITY}
  \hfill \textit{The indulgence is a single bell-note, held in a jar. Break the jar to release it.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:mistlands-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (passes/seals/tokens) that change position rather than call for a roll.
\item If any A appears, echo \textbf{breath and boundary} --- bells answering across water, salt that won't dry, voices returning late.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:mistlands-mechanics}

\begin{itemize}
\item \textbf{Breath-Toll:} Many claim the mist remembers every breath. Travelers hold breath crossing bridges, fearing the mist will count them wrong. Mechanically, when you cross a threshold without a ward-salt or bell, roll Spirit + Resolve (DV~$3$). On a failure, the mist learns your name --- gain the \texttt{MARKED} condition (the next spirit encounter starts at Desperate Position).
\item \textbf{Salt as Promise:} A handful of salt is more than warding --- it is a vow. Spill it on plank or stone to swear safety. Once per session, a character may spend a handful of ward-salt to create a \textbf{Salt Oath}~[$4$] timer. While the timer has segments, the oath-taker gains $+1$ die to resist fear or possession. If the oath is broken, the timer fills and the breaker takes Harm~$1$ (spiritual backlash).
\item \textbf{The Ninth Bell:} Legends say a ninth bell once rang on the Line. The note is not in any ringbook; no warden speaks of it. Some nights, though, it rings --- and gates open where none should be. When the GM spends $3$+~SB in a scene, they may declare that the Ninth Bell rings. All characters must test Spirit + Resolve (DV~$4$) or lose $1$ Fatigue (the sound steals breath).
\item \textbf{Payden's Port:} The only true city in the Mistlands, huddled on the western coast of the Aberderrin. Its wharves are built from salvaged longship timber, and its law is split between the Aeler Legate's seal and the old bell-wardens' traditions. Grain ships leave from its harbour each autumn, carrying the Protectorate's due east --- and returning with iron, engineers, and the occasional missionary. The port's inn, the \emph{Drowned Rat}, has a bell that rings once for every sailor lost at sea. It rings often.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Bell-Keeper} --- The Sealed Gate. \textit{A bell that never rings, yet holds the threshold closed.}
    \item \textbf{The Salted One} --- Oath of Mercy and Grace. \textit{A prayer that preserves body and soul, but remembers the name of everyone it saves.}
    \item \textbf{The Mist-Father} --- Khemesh. \textit{The pressure beneath the fog; the hunger that waits in still air.}
    \item \textbf{The Lantern Bearer} --- The Traveler. \textit{The light that guides lost ships --- and lost souls.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{Saint Lucas of the Iron Bell (The Sealed Gate):} The bell-wardens venerate a legendary warden who hammered a ninth bell into the Weeping Gate to seal a Direwood breach. The bell cracked, but the breach closed. His chisel is kept at Payden's Port; it is said to quench the mist for an hour if struck against an iron post. The Aeler Protectorate has tried to confiscate it twice; both attempts ended with the confiscators lost in fog.
  \item \textbf{Saint Anja of the White Pans (Oath of Mercy and Grace):} A salt-monk who died blessing the pans during a tide-mist. Her body never decayed, and her chapel now serves as a hospice for fog-sick sailors. The Temple of Light sends offerings of fresh water each spring; the Everflame considers her a heretic. Pilgrims who pray at her shrine may ask one question of a drowned ancestor --- but the answer is always a single word, and that word is never given freely.
  \item \textbf{The Grey Hospice of the Unnamed:} Sister Darya runs a hospice near Payden's Port. She asks no payment, but she records every patient's name in a book that never dries. The Aeler clerks suspect she is erasing names, not healing wounds. She has been investigated three times; each time, the investigator vanished into the fog and returned with no memory of why they came.
  \item \textbf{The Bell-Ringer's Covenant (The Keeper of the Silent Seal):} Notary Voss is the Protectorate's appointed judge, charged with imposing Aeler law on local customs. He carries a water-timer to measure the shifting of legal moods, but the mist ignores it. His rulings are always delivered in triplicate, and the bell-wardens obey them --- slowly. Those who lie before him find that their next passage through the Ghost-Ferry Slip is denied; the ferryman refuses to name them.
  \item \textbf{The Ninth:} In any Mistlands scene, the ninth person to speak in a room draws the mist's attention. The wise hold their tongues until someone else breaks the count. Those who speak ninth are said to see a pale figure standing at the foot of the bed that night --- waiting to take a breath.
  \item \textbf{The Black Barrow Keep (Aelinnel Ruin):} Looming over Payden's Port, this ancient workshop of bell-automata and ward-forges is sealed by the Protectorate --- but smugglers and treasure-hunters still enter. Locals whisper that the lowest catacombs connect to the Ways Between. The Aeler have tried to collapse the upper levels three times; each time, the rubble was found cleared by morning, with no tracks in the fog.
\end{itemize}

\subsection*{Mistlands Superstitions (burn for $+1$ die once)}
\begin{itemize}
  \item \textbf{Bread to the Fog:} Toss a crust into the mist before crossing a bridge. Ignore the first \emph{Ground-Mist} penalty. \card{\clubsuit}{SUPERSTITION}
  \item \textbf{Hold the Ninth Breath:} Cross a threshold while holding your breath. Cancel \emph{Witchlight Count} or take $-1$ die on your next action (your choice). \card{\clubsuit}{SUPERSTITION}
  \item \textbf{Name to the Salt:} Whisper a dead enemy's name into a salt pan. Gain $+1$ Effect against undead for the scene, but mark \emph{Obligation} to the Legate. \card{\clubsuit}{SUPERSTITION}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:mistlands-sb}

\paragraph{Minor Complications ($1$~SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{wardens or fog-knights}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{spirits or patrol boats}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{trackers or mist-spirits}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}.
\item \textbf{Supply Strain:} Mark $+1$ segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks ($2$~SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Protectorate forces or local warden} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{mist shift or bell failure}.
\item \textbf{Foe Appears:} A \textbf{wraith, raiding party, or inspector} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble ($3$~SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{fog-knights, spirits, or Aeler forces} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes --- \textbf{bell-line fails/ward broken/mist claims name}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by $1$ segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue~$1$/Harm~$1$/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns ($4$+~SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Legate, Lord Warden, or Mist-Seer} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically --- \textbf{mist thickens/bells ring/forest breathes}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice --- \textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Mistlands (Bells \& Breath):} Bells ring without cause, breath forms visible words, mirrors reflect wrong faces.
\item \textbf{Mistlands (Salt \& Warding):} Salt refuses to dry, ward-runes bleed color, protection tokens crack.
\item \textbf{Mistlands (Direwood Threat):} Trees whisper your name, roots shift position, forest breath grows stronger.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Navigation):} On a hit, shift one band. On a strong hit, also impose \emph{Fog Bank}.
\item \textbf{Silence the Bell (Presence + Stealth):} On a hit, freeze range for one exchange; if a ward-salt pouch is in your possession, $+1$ die.
\item \textbf{Weather the Tide (Resolve + Survival):} On a hit in \emph{Tide-Mist}, you pick who keeps their bearings; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Bell \& Breath)}
\begin{itemize}
\item \textbf{The Ninth Bell Ringer (Direwood herald)} --- \emph{Phases}: Lost Names $\rightarrow$ Broken Lines $\rightarrow$ Open Gate. \emph{Cracks}: restore a ringbook; trade a breath; rehang a forbidden bell with copper nails.
\item \textbf{Queen of Counting Lights (fae arbiter)} --- \emph{Phases}: Courtesies $\rightarrow$ Geas-Market $\rightarrow$ Counting of Leaves. \emph{Cracks}: return stolen way-marks; speak said and meant; gild the iron.
\item \textbf{Reed-Wolf Captain (Linn raider)} --- \emph{Phases}: Smoke Signals $\rightarrow$ Night Oars $\rightarrow$ Thing Verdict. \emph{Cracks}: name a feud-token; win a skald-duel; force harbour-truce under bell and banner.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In the Mistlands, power flows through ritual, warding, and the careful balance between the living and the dead. The Aeler Protectorate maintains nominal control through military presence and legal frameworks, but true authority lies with those who understand the bells, the salt, and the breath of the land itself. Local wardens, salt-monks, and bell-keepers hold sway over daily life, while the Legate of the Mists mediates between imperial law and ancient custom.

\textbf{For the GM:}  
Patronage in the Mistlands revolves around protection --- from mist, from spirits, from the encroaching forest. Rewards often take the form of tokens, seals, or safe passage that can be turned into greater shelter. To emphasize the folk-horror feel:
\begin{itemize}
\item Tie rewards to visible symbols (salt, bells, lanterns) that can be challenged, depleted, or voided by the mist itself.
\item Let rival factions (Protectorate vs. bell-wardens vs. Direwood refugees) issue conflicting protections, forcing players to choose whose favor matters more.
\item Use the levees, ferries, and watchtowers as arenas for social contests, where knowledge of the old ways (counting breaths, naming spirits, offering salt) determines success.
\end{itemize}
In the Mistlands, your safety is your standing, and your standing depends on who remembers your name.
\end{tcolorbox}

% Index entries
\index{Mistlands|see{Bells, Salt, and Breath}}
\index{Theme generator}
\index{Bells, Salt, and Breath generator}
\index{Places!Mistlands}
\index{People!Mistlands}
\index{Complications!Mistlands}
\index{Rewards!Mistlands}
\index{Aeler!Mistlands}
\index{Protectorate!Mistlands}
\index{Weeping Gate}
\index{Witchlight Bridge}
\index{Bell-Line levee}
\index{Ghost-ferry}
\index{Pall Watch-tower}
\index{High-Mist Pass}
\index{Legate of the Mists}
\index{Lord Warden of the Fens}
\index{Mist-Seer}
\index{Fog-knight}
\index{Oath-ferryman}
\index{Bell-warden}
\index{Salt-monk}
\index{Shroud-diver}
\index{Ward-salt}
\index{Pall Indulgence}
\index{Wraith-pardon}
\index{Direwood}
\index{Valewood}
\index{Payden's Port}
\index{Black Barrow Keep}
\index{Saint Lucas of the Iron Bell}
\index{Saint Anja of the White Pans}
\index{Grey Hospice of the Unnamed}
\index{Bell-Ringer's Covenant}
\index{Local Patrons!Mistlands}

\subsection*{Black Barrow Keep --- The Ruin Above Payden's Port}
\label{sec:black-barrow}

Looming over the eastern edge of Payden's Port, Black Barrow Keep is a jagged wound in the skyline. Its basalt walls weep salt, and its broken towers lean into the fog like old men listening for a bell that never rings. The Aelinnel built it in a forgotten age---not as a fortress, but as a \emph{workshop}. Here, they forged bell-automata, copper sentinels, and the first ward-lines that still keep the Direwood at bay. When the Mistlands became a Protectorate, the Aelinnel sealed the deep forges and retreated. The locals say the keep is haunted. The truth is worse: it is not haunted---it is \emph{occupied}.

Below the keep, the catacombs stretch down through nine known levels. Each level was sealed by a different generation of wrights: Aeler engineers, Aelinnel bell-smiths, Linn shipwrights, and older crafts whose names are scrubbed from the records. The lowest level is said to contain a working that has been ticking since before the Hollow learned to count.

\textbf{Accessing the Keep:} The main gate is chained with a bell-lock that requires a copper key and the correct sequence of rings (eight, never nine). Side entrances include a collapsed ropeway from the salt pans, a flooded stair from the Ghost-Ferry Slip, and a smugglers' tunnel that exits near the Grey Hospice. The Protectorate has officially sealed the keep, but wardens look the other way for a share of salvage---or a story about what moves in the deep.


\begin{tcolorbox}[
  enhanced,
  breakable,
  colback=black!3,
  colframe=blue!60!black,
  title={Generator --- Black Barrow Keep: Rooms \& Encounters},
  left=2mm,right=2mm,top=2mm,bottom=2mm
]
\medskip

\begin{longtable}{|c|p{0.75\linewidth}|}
\hline
\textbf{d12} & \textbf{Area / Encounter} \\ \hline
\endfirsthead
\hline
\textbf{d12} & \textbf{Area / Encounter} \\ \hline
\endhead

1 & \textbf{The Hanging Bell-Garden.}
    A collapsed floor reveals a chamber where bells grow from the
    ceiling like stalactites. Each bell has a clapper shaped like a
    finger bone. \textit{Effect: A bell that rings once for each secret
    spoken beneath it. The ringing can be heard throughout the keep.} \\ \hline

2 & \textbf{The Copper Forge.}
    Still warm, though the fire went out centuries ago. A half-completed
    automaton sits on the anvil. Its brass skull is hinged open; inside,
    a nest of salt-eels writhes. \textit{Effect: The forge can be
    reactivated with a bell-key, producing a single ward-salt token per
    scene.} \\ \hline

3 & \textbf{The Oubliette of the Unnamed.}
    A pit lined with counting marks --- eight scratches, a gap, eight
    scratches. At the bottom, a wright's skeleton still holds a
    hammer. The hammer rings when you touch it.
    \textit{Effect: The hammer can drive a copper nail into any surface,
    sealing a threshold for one scene.} \\ \hline

4 & \textbf{The Echo Scriptorium.}
    Shelves of slate tablets, each inscribed with a single number. The
    numbers are not in sequence. A ghost wright sits at a desk,
    writing the same number over and over.
    \textit{Effect: The ghost will answer one question about the keep's
    history, but will then ask for a number you have never written.} \\ \hline

5 & \textbf{The Drowned Armoury.}
    Sealed by a slow leak from the salt pans. Weapons float in brine.
    One of them --- a bell-hilted sword --- hums when you near it.
    \textit{Effect: The sword counts your heartbeats. If you draw it in
    combat, you gain $+1$ die, but the sword will demand a name at the end
    of the scene.} \\ \hline

6 & \textbf{The Weeping Cistern.}
    A stone basin filled with black water. The water rises when you
    speak. A voice from below asks, ``What is the ninth number?''
    \textit{Effect: Answer with silence, and the water recedes,
    revealing a sealed door. Answer incorrectly, and the cistern floods
    the room.} \\ \hline

7 & \textbf{The Cogwheel Choir.}
    A chamber of interlocking copper gears, each tooth stamped with a
    rune. When the wind blows through the broken wall, the gears turn
    and the runes sing a cracked hymn.
    \textit{Effect: A character who learns the tune may reroll one
    failed navigation test while in the Mistlands (once).} \\ \hline

8 & \textbf{The Salt-Liche's Studio.}
    A wright who refused to die. Its flesh is replaced by
    salt-cured leather and copper wire. It still tries to finish a bell
    that has no clapper.
    \textit{Effect: The salt-liche is not hostile, but will trade a
    bell-key for a memory of fresh water.} \\ \hline

9 & \textbf{The Unstable Lift.}
    A chain-hoist that descends into darkness. The chains are warm.
    The lift stops at every eighth floor, never the ninth.
    \textit{Effect: Riding the lift skips two levels of the catacombs,
    but ticks the \textsc{Fog Attention} timer by $1$.} \\ \hline

10 & \textbf{The Bone Gallery.}
    A corridor lined with the remains of wrights from every generation.
    Each skeleton wears a different style of bell-maker's apron. The
    newest apron is still damp.
    \textit{Effect: A PC may spend a scene searching; on a
    Wits+Lore roll (DV~$4$), they find a key, a token, or a warning.} \\ \hline

11 & \textbf{The Ticking Heart.}
    A sealed chamber behind a copper door. Something inside ticks at a
    steady rhythm. The rhythm changes when you lie.
    \textit{Effect: The door requires a blood offering and a true
    confession to open. Inside is a working that can reset one broken
    bell-line --- but it will ask for a year of your life.} \\ \hline

12 & \textbf{The Ninth Gate.}
    A door that appears only when the keep's bells have rung eight
    times in sequence. Beyond it, a spiral stair descends into absolute
    silence. No echo returns.
    \textit{Effect: This leads to the lowest catacomb, where the first
    wright's original bell-line generator still turns. Activating it
    would drive the Direwood back a mile --- but the noise would wake the
    Ninth Bell Ringer.} \\ \hline
\end{longtable}

\bigskip
\textit{To use:} Roll once for the room, then optionally roll again for
a resident, hazard, or treasure from the appropriate tables in
\S\ref{sec:mistlands-places}, \S\ref{sec:mistlands-people}, or
\S\ref{sec:mistlands-complications}. For the deeper catacombs, increase
the DV of all spirit-related rolls by $1$ and add a \textsc{Stale Air}
timer~[$4$] at the start of each scene.
\end{tcolorbox}

\paragraph*{Factions in the Keep}
\begin{itemize}
\item \textbf{The Bell-Wrights' Ghosts:} Ancient Aelinnel spirits who still maintain the lower forges. They ignore the living unless the forges are disturbed. Offerings of copper nails or a correctly rung bell will buy safe passage.
\item \textbf{The Protectorate Salvage Team:} Aeler engineers and a squad of fog-knights, ostensibly ``surveying'' the keep for hazards. They will confiscate any working bell they find and arrest anyone without a Protectorate writ. Their captain has a brass hand that is also a tuning fork.
\item \textbf{The Hollow's Anticipators:} A cult that believes the keep was built to contain the Hollow's attention. They leave offerings of salt and silence. They are not dangerous --- but they will follow you, writing down every bell you ring and every name you speak.
\end{itemize}

\paragraph*{Catacombs Depth Timer }
For expeditions that descend below the first sub-level, track a \textbf{Depth Timer [6]}. Each time the party descends a stair, uses the unstable lift, or passes a sealed door, tick the timer. At 3 segments: all navigation rolls gain $+1$ DV (the walls shift). At 6 segments: the party reaches a sealed level that requires a bell-key, a blood offering, or a true name to open. The deepest level (Depth 6 or more) connects to the Ways Between --- a mistake may drop the party into the Hollow's own threshold.