\section{The Mountain Satrapies --- ``The Iron Passes''}
\label{chap:satrapies}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The mountain weighs every breath, every ingot, every oath. Pay in iron, pay in silence, or pay in blood.}

\begin{quote}
  \textbf{Viceroy Thrain (Elite):}  
  ``The passes do not care for your lineage or your gods. They care for iron, tea, and the weight of a sworn ledger. Aeler and Dhaharan have bled into the same stone here. I wear a dwarf's breath-counter and a Dhaharan's fire-token. The mountain accepts both.''
\end{quote}

\begin{quote}
  \textbf{Elephant Mahout (Commoner):}  
  ``My grandfather drove elephants for the Sun Kings. My father drove them for the Aeler. I drive them for the Viceroy. The elephants do not care who holds the howdah. They care that the pass is clear and the fodder is fresh.''
\end{quote}

\textbf{Genre / Mood:} High‑altitude contract horror, industrial colonialism, elephant‑driven logistics.

\textbf{Starting Location:} The Viceroy's Hall, a longhouse of timber and stone, with a throne of iron ore that no one sits on. Viceroy Thrain stands behind it, a bell around his neck: ``The lease is up for renewal. The rajas want lower tithes. The Admiralty wants the passes opened to their trade. And the mines are whispering of a demon waking. You are not here by accident. Tell me whose interests you serve — and whose name I should carve on the next treaty.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Mountain Satrapies are the northern highlands of Dhahara, where the Aeler viceroys rule on behalf of the distant High King. Here, the air is thin and cold, the rivers run white with snowmelt, and the passes are measured in breath-counts rather than miles. Iron mines scar the hillsides, worked by Dhaharan labourers and overseen by Aeler engineers. Elephant caravans haul ore down to the Brass Coast; mule trains carry silver and tea up to the mountain holds. The Aeler have established a viceroyalty here — not conquest, but a long lease. The local rajas pay tribute in iron and receive protection from lowland enemies. It is an uneasy peace, written in bell-codes and toll-ledgers. The architecture blends Aeler stonework with Dhaharan domes; the cuisine mixes dwarf-bread with spiced tea; the law is a creole of Covenant statutes and Aeler breath-bonds.

\vspace{2mm}
\textbf{What you notice first:} The ringing of iron hammers and the lowing of elephants. The way every bridge has a dwarf toll-keeper with a brass scale. The silence after a bell rings — the silence is for the mountain to answer.

\textbf{The one rule that kills newcomers:} Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one — a breath, a memory, a shadow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Lease}

Power here is measured in iron quotas, tea taxes, and the length of a viceroy's lease. The Satrapies are not a colony — the Aeler High King holds the passes by treaty, renewed every nine years. The Viceroy answers to the High King, but also to the local rajas, the Covenant judges, and the elephant clans. Every decision is a negotiation between mountain custom and lowland law. The Aeler engineers speak a creole of their own tongue; the Dhaharan nobles wear Aeler breath-counters as jewellery; and the elephants — the elephants remember who fed them during the famine of the ninth lease.

\subsection*{Prominent NPCs of the Mountain Satrapies}
\label{sec:satrapies-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Viceroy Thrain & Aeler governor of the Iron Passes & Maximise ore output for the High King & He wears a bell around his neck; it rings when he lies. \\
\hline
Rani Jaya & Local raja of the eastern hills & Balance tribute to Aeler with loyalty to her clan & She lost an arm to a mine collapse; the prosthetic is a pickaxe. \\
\hline
Mine-Captain Korr & Aeler engineer & Keep the tunnels safe and the quotas met & He counts his breaths in eights; he has never reached nine. \\
\hline
Tarkan & Elephant caravan master & Move iron to the coast before the snows close the passes & He speaks to his elephants in Dhaharan; they answer in Aeler. \\
\hline
Sister Gaura & Covenant judge of the hill circuit & Resolve disputes between Aeler law and local custom & She carries a water-clock that runs on tears — her own. \\
\hline
\end{longtable}

\subsection*{Names of the Mountain Satrapies}
\label{sec:satrapies-names}

A mix of Aeler gutturals and Dhaharan vowels. Surnames may be Aeler clan names or Dhaharan hill titles.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/House} & \textbf{Epithets} \\
\hline
Thrain & Jaya & Ironfoot & \textit{the Bell-Breaker} \\
\hline
Korr & Gaura & Stonehand & \textit{Breath-Counter} \\
\hline
Borin & Meera & Vikram & \textit{the Elephant's Friend} \\
\hline
Halk & Sita & Aezelbron & \textit{Bridge-Warden} \\
\hline
Rurik & Laxmi & Thrainsdottir & \textit{Toll-Eater} \\
\hline
\end{longtable}

\paragraph*{(Mine/Bridge/Hall)} Iron mine with a carved face above the entrance; Aeler toll bridge with a dwarf and a scale; Viceroy's hall with a throne of iron ore.

\section*{Spades --- Places}
\label{sec:satrapies-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Iron Mine of the Crying Face} --- A deep shaft with a carved face above the entrance.
  \hfill \textit{The face weeps rust. The tears are iron oxide. The mountain is mourning its dead.} `[BREATH]`
\item \textbf{Aeler Toll Bridge} --- A stone arch over a gorge; a dwarf with a scale sits at the far end.
  \hfill \textit{The toll is not in coin — it is in weight. Your truth is weighed against a bar of iron.} `[TOLL]`
\item \textbf{Viceroy's Hall} --- A longhouse of timber and stone; a throne of iron ore.
  \hfill \textit{The throne is cold. The viceroy never sits on it — only stands behind it.} `[LEASE]`
\item \textbf{Elephant Corral (Mountain)} --- A fenced area with heavy chains; elephants are chained by one leg.
  \hfill \textit{The chains have names — each is the elephant's first kill.} `[COMMUNITY]`
\item \textbf{Silent Pass} --- A high mountain pass where no bird flies.
  \hfill \textit{The air is too thin for sound. Arguments must be written in the snow.} `[BREATH]`
\item \textbf{Mine of the Ninth Seal} --- A sealed tunnel with a Covenant seal on the timbers.
  \hfill \textit{Inside is a demon that the miners woke. The seal is cracking.} `[MAGIC]` `[LEASE]`
\item \textbf{Tea Terraces (Aeler-Dhaharan)} --- Stone-step fields where tea shrubs grow.
  \hfill \textit{Each terrace is a different clan's share of the harvest. The tea tastes of slate.} `[TRADE]`
\item \textbf{Dwarf Counting-House} --- A low building with brass scales and an iron ledger.
  \hfill \textit{The ledger records breath‑bonds. It weighs the air.} `[TOLL]`
\item \textbf{Raja's Fort (Hill)} --- A stone fortress with painted gates and a single tower.
  \hfill \textit{The tower is also a fire shrine. The flame is lit with dwarf-coal.} `[COMMUNITY]`
\item[J] \textbf{Bridge of Nine Echoes} --- A rope bridge over a chasm; each step echoes nine times.
  \hfill \textit{The ninth echo is always a whisper. It is the name of the next person to fall.} `[OMEN]`
\item[Q] \textbf{Abandoned Smelter} --- A ruined furnace where iron was once refined.
  \hfill \textit{The slag heaps are still warm. Something lives in the warmth.} `[MAGIC]`
\item[K] \textbf{Viceroy's Archive} --- A subterranean vault of treaties and toll receipts.
  \hfill \textit{The archive is sealed with Aeler keystone wax and Dhaharan clay.} `[LEASE]`
\item[A] \textbf{The First Lease Stone} --- A monolith carved with the original treaty between Aeler and Dhahara.
  \hfill \textit{The stone is warm. It glows when the lease is renewed. It cracks when the lease is broken.} `[LEASE]` `[OMEN]`
\end{enumerate}

\paragraph*{(Engineer/Raja/Judge)} Mine-captain with a tuning fork and a level; Rani Jaya with a pickaxe prosthetic; Sister Gaura with a water-clock that runs on tears.

\section*{Hearts --- People \& Factions}
\label{sec:satrapies-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Mine-Captain (Aeler)} --- A dwarf with a tuning fork and a level. He can hear a crack in a beam from a hundred paces.
  \hfill \textit{He never walks under a keystone. The mountain has ears.} `[BREATH]`
\item \textbf{Rani Jaya} --- A young woman in armour, riding a caparisoned elephant. She lost an arm to a mine collapse.
  \hfill \textit{The prosthetic is a pickaxe. She can write with it. The letters are iron filings.} `[COMMUNITY]`
\item \textbf{Covenant Judge (Sister Gaura)} --- Robed in grey, a water-clock on her belt. The clock runs on tears.
  \hfill \textit{She refills it from her own eyes when she hands down a death sentence.} `[LEASE]`
\item \textbf{Elephant Caravan Master (Tarkan)} --- Old, missing two fingers, tattooed with elephant footprints.
  \hfill \textit{He knows the name of every elephant that has crossed the pass in fifty years.} `[COMMUNITY]`
\item \textbf{Aeler Toll-Keeper} --- A dwarf with a brass scale and a ledger. He weighs your goods and your lies.
  \hfill \textit{The scale tips for both. A lie makes the pan dip lower.} `[TOLL]`
\item \textbf{Smelter Worker} --- A young man with red-rimmed eyes. He coughs iron dust.
  \hfill \textit{His spittle stains the ground. He will not live past thirty. He knows.} `[LABOR]`
\item \textbf{Treasure Hunter} --- A foreigner digging in an old landslide. He is looking for the tomb of a prince.
  \hfill \textit{The prince is the one who stole the Kalrantha amulet. The treasure hunter is a fool.} `[MAGIC]`
\item \textbf{Raja's Spymaster} --- A woman with a limp; she uses a walking stick that is also a blowpipe.
  \hfill \textit{The darts are tipped with tea — caffeine that stops hearts.} `[POLITICS]`
\item \textbf{Tea Trader (Fhara)} --- A merchant with a caravan of mules. He sells tea to the Aeler; they pay in iron.
  \hfill \textit{He sells iron to the coast; they pay in silver. He is the richest man in the passes.} `[TRADE]`
\item[J] \textbf{Blind Pilgrim} --- A woman in white, walking the passes alone. She is searching for a shrine where a bell rang without a clapper.
  \hfill \textit{The shrine is a cave. The cave is the mine of the ninth seal.} `[FAITH]`
\item[Q] \textbf{Ghost of the Landslide} --- A figure in grey, seen only at dusk. It points at the pass.
  \hfill \textit{It is pointing at a safe route — or a trap. No one has survived to find out.} `[OMEN]`
\item[K] \textbf{Viceroy's Cipher Clerk} --- A dwarf who writes diplomatic letters in code.
  \hfill \textit{The code is a poem. The poem is about a lost mine. The mine is real.} `[LEASE]`
\item[A] \textbf{Demon's Whisper (from the Ninth Seal)} --- A voice heard in the sealed mine.
  \hfill \textit{It offers a deal: a year of good fortune for a single memory. The memory is the last breath of a dying friend.} `[MAGIC]` `[LEASE]`
\end{enumerate}

\paragraph*{(Cave-in/Toll/Echo)} A cave-in traps miners; the Aeler toll-keeper demands double payment; the ninth echo on the bridge speaks a name.

\section*{Clubs --- Complications/Threats}
\label{sec:satrapies-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Cave-In (Sabotage)} --- A tunnel collapses. The collapse was not an accident — a saboteur loosened the timbers. `[DISASTER]`
  \hfill \textit{The saboteur is in the party. Or the party is accused.}
\item \textbf{Toll Dispute} --- The Aeler keeper demands double; the raja's men refuse. The dispute is a trap. `[TOLL]` `[POLITICS]`
  \hfill \textit{Both sides want the party to choose a side. Neither is honest.}
\item \textbf{Ninth Echo} --- On the Bridge of Nine Echoes, someone counts the echoes. The ninth echo says a name. `[OMEN]` `[FEAR]`
  \hfill \textit{The name is the speaker's. They faint. The bridge remembers.}
\item \textbf{Landslide} --- The pass is blocked. The alternative route is three days longer. `[WEATHER]` `[TRAVEL]`
  \hfill \textit{The landslide was deliberate — a rival caravan master wants to delay you.}
\item \textbf{Breath Shortage} --- At high altitude, the party struggles. Mark Fatigue 1 per scene. `[BREATH]` `[ENDURANCE]`
  \hfill \textit{Acclimatisation requires a rest day. The rest day will cost a toll.}
\item \textbf{Spy in Camp} --- A worker is stealing iron samples. The spy is working for an Ashaani agent. `[POLITICS]` `[STEALTH]`
  \hfill \textit{The samples are for a new alloy. The alloy is a weapon.}
\item \textbf{Covenant Audit} --- Sister Gaura demands to inspect the party's cargo. `[LEASE]` `[JUSTICE]`
  \hfill \textit{The cargo includes a relic from the Red Desert — a stolen Covenant seal.}
\item \textbf{Aeler Bell Code Change} --- The toll-keeper rings a different sequence; the party is now ``unregistered.'' `[TOLL]` `[AUTHORITY]`
  \hfill \textit{The change was ordered by the viceroy. The viceroy is testing your loyalty.}
\item \textbf{Elephant Sickness (Curse)} --- A disease spreads through the caravan. The disease is a curse from a local spirit. `[COMMUNITY]` `[MAGIC]`
  \hfill \textit{The spirit was disturbed by the mine's expansion. It wants a sacrifice — a memory, not blood.}
\item[J] \textbf{Viceroy's Summons} --- Thrain demands the party attend him. He offers a deal: find the saboteur in exchange for a toll waiver. `[LEASE]` `[POLITICS]`
  \hfill \textit{The saboteur is his own clerk. The clerk is his daughter.}
\item[Q] \textbf{Silent Pass's Cold} --- The temperature drops; the party loses 1 Supply per hour. `[WEATHER]` `[BREATH]`
  \hfill \textit{The cold is supernatural — a frost spirit has taken residence. It wants a memory.}
\item[K] \textbf{Demon Wakes (Ninth Seal)} --- The sealed mine cracks open. A shadow escapes; it possesses the nearest Aeler engineer. `[MAGIC]` `[LEASE]`
  \hfill \textit{The engineer begins to sabotage the pass. The demon wants the lease broken.}
\item[A] \textbf{Lease Stone Cracks} --- The First Lease Stone glows red and splits. The treaty is void. War is coming. `[LEASE]` `[APOCALYPSE]`
  \hfill \textit{The party has one day to flee the passes before the Aeler and Dhaharan armies clash.}
\end{enumerate}

\paragraph*{(Token/Ingot/Seal)} Aeler breath-counter (silver bell, wear to count breaths safely); iron ingot (stamped with the viceroy's mark); Covenant seal (clay, for witnessing oaths in the passes).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:satrapies-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Aeler Breath-Counter (Silver Bell)} --- Wear it to count your breaths in high altitude. The bell rings when you reach eight. `[BREATH]` `[TRAVEL]`
  \hfill \textit{Stop at eight. The ninth breath is the mountain's.}
\item \textbf{Iron Ingot (Viceroy's Mark)} --- A single bar of high-grade iron, stamped with a dragon. Worth 1 Measure. `[TRADE]` `[LEVERAGE]`
  \hfill \textit{The dragon is the viceroy's mark. Present it to any Aeler forge for a favour.}
\item \textbf{Covenant Seal (Clay, Himadri)} --- A lump of clay. Heat it to witness any oath in the foothills. `[LEASE]` `[JUSTICE]`
  \hfill \textit{The heat comes from a coal taken from the sealed mine. The coal whispers.}
\item \textbf{Toll Waiver (Brass Nail)} --- A brass nail. Present it to pass any Aeler bridge for free. `[TOLL]` `[TRAVEL]`
  \hfill \textit{The nail fits into a hole in the toll-keeper's scale. The hole is for bribes.}
\item \textbf{Elephant Whistle (Bone)} --- A bone whistle. Blow it to call an elephant from the caravan. `[COMMUNITY]` `[TRAVEL]`
  \hfill \textit{The whistle is made from the tusk of a bull that died in a rockslide. Its ghost runs with the elephant.}
\item \textbf{Raja's Signet Ring (Silver)} --- A silver ring with a tiger. Commands loyalty from the raja's clan. `[COMMUNITY]` `[SOCIAL]`
  \hfill \textit{The ring is a copy. The real one was lost in a landslide. The ghost of the original owner wears it.}
\item \textbf{Smelter's Gloves (Asbestos)} --- Handle hot metal safely. The gloves are cursed: they itch when you lie. `[CRAFT]` `[TRUTH]`
  \hfill \textit{The itch is mild at first. It becomes unbearable with repeated lies.}
\item \textbf{Map of the Silent Pass (Leather)} --- A chart drawn on leather. Shows a safe route through the high pass. `[TRAVEL]` `[SECRET]`
  \hfill \textit{The map is incomplete. The missing section is a fold. The fold is a secret — a cave with a spring.}
\item \textbf{Viceroy's Letter of Introduction} --- A sealed scroll. Grants access to any Aeler hold in the mountains. `[LEASE]` `[AUTHORITY]`
  \hfill \textit{The letter is written in blood. The blood is the viceroy's. He changes his blood type each season.}
\item[J] \textbf{Blind Pilgrim's Staff} --- A wooden staff with a bell on top. Ring it to summon a guide. `[FAITH]` `[TRAVEL]`
  \hfill \textit{The guide is a ghost — the pilgrim's predecessor. She will lead you safely, but she will ask for a story.}
\item[Q] \textbf{Demon's Nail (Sealed Mine Timber)} --- A piece of the sealed mine's timber. Touch it to a lock; the lock opens. `[MAGIC]` `[STEALTH]`
  \hfill \textit{The lock also screams. The scream alerts the demon.}
\item[K] \textbf{Ninth Echo's Whisper (Recording)} --- A piece of slate carved with the ninth echo's words. Play it to silence a foe. `[COMBAT]` `[FEAR]`
  \hfill \textit{The foe hears their own name and freezes for one exchange. The slate then crumbles.}
\item[A] \textbf{Viceroy's Forgiveness (Pardon)} --- A written pardon for any crime in the Satrapies. `[LEASE]` `[AUTHORITY]`
  \hfill \textit{The pardon is conditional. The condition is a secret mission. The mission is to assassinate a rival viceroy.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:satrapies-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2–5 → 4, 6–10 → 6, J/Q/K → 8, A → 10).
\item Diamonds are codified outcomes (bells, ingots, nails) that change position rather than call for a roll.
\item If any A appears, echo \textbf{mountain-omens}—iron that weeps rust, bells that ring without clappers, a pass that closes behind you.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:satrapies-mechanics}

\begin{itemize}
\item \textbf{The Ninth Breath:} Aeler miners believe that counting to nine in a mine summons the mountain's attention. They count in eights, resetting after each set. Visitors who count to nine aloud may trigger a cave-in — or worse. Mechanically, any character who voluntarily counts to nine in a mine must test Resolve (DV 4) or mark 1 Fatigue as the mountain takes a breath. On a critical failure, the mountain takes a memory instead (lose 1 XP or forget a minor detail).
\item \textbf{Lease Clock [8]:} Tracks the political pressure around the treaty renewal. The clock ticks when the party assists or hinders a faction (Aeler, rajas, Admiralty, Covenant). At 4 segments: a faction threatens to walk out of negotiations. At 8: the lease is broken or renewed in the party's favour (or against it). The party can reset the clock by brokering a concession.
\item \textbf{Elephant Community Share:} Instead of individual ownership, elephant caravans operate on a share system. A character who helps the caravan (clearing a pass, feeding a sick elephant, healing the mahout) gains a \textbf{Share Token}. Spend it to call on the caravan for a single favour: passage for a wagon, a day's fodder, or a warning about a rival. Share Tokens are not debt — they are reciprocity.
\item \textbf{Breath-Bonds as Leverage:} Aeler engineers use breath-bonds to secure contracts. A character who holds a breath-bond can demand a renegotiation of tolls, quotas, or labour terms. The bond is not a debt — it is a recorded agreement that both parties benefit from keeping. Breaking a breath-bond does not incur a curse; it simply ends the trust, and the other party will never deal with you again.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Thrain the Counter} — Mykkiel. *The divine auditor; every deed must balance.*
    \item \textbf{The Silent Bell} — The Sealed Gate. *The threshold that asks only that you stop at eight.*
    \item \textbf{The Forge Father} — Adur (Oath of Flame and Light). *The flame that tempers iron and burns lies.*
    \item \textbf{The Old Breath} — The Traveler. *The wind that carries caravans and counts their steps.*
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Ninth Seal Mine and the Kalrantha Amulet (Dhahara):} The demon sealed in the mine is a fragment of Varash the Unmaker. The Veiled Sisters want to free it to use against the Covenant. The Breath-less want to re-seal it. The demon speaks in the voices of dead miners.
  \item \textbf{Iron Trade and the Ngomebe (Stone Men):} The iron from the Satrapies is used to forge rollers for the walking cities of Ngomebe. A disruption in the iron supply could stop a city mid-migration. The Mason-Kin have a permanent envoy at the Viceroy's hall.
  \item \textbf{Covenant Judges and the Sidhi (Way of Silk):} Sister Gaura is a Covenant judge; her water-clock is a Sidhi design. She may have contacts in the Brass Coast who can verify a freedman's seal. She also knows which freedmen are still being hunted.
  \item \textbf{The Blind Pilgrim and the Hollow (Eastern Realms):} The pilgrim is searching for a shrine that marks the Hollow's boundary in the high passes. Aelaerem hedge-keepers know the location but refuse to speak it. They say the shrine is a missing ninth step in the mountain's own counting song.
\end{itemize}

\subsection*{Mountain Satrapies Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Mine:} Crumble a crust into the tunnel mouth. Ignore the first \emph{Cave-In} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Bell Rope:} Tie a knot in the rope of a shrine bell. Cancel \emph{Ninth Echo} or take –1 die on your next stealth roll (the bell still rings, but muffled). `[SUPERSTITION]`
  \item \textbf{Name to the Iron:} Whisper a dead miner's name into an ingot. Gain +1 Effect when negotiating with Aeler this scene, but mark \emph{Obligation} to the Viceroy. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:satrapies-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Aeler engineers or raja's spies}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{mine guards or toll-keepers}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{elemental spirits or rival caravans}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favourable toll rates}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local mine-captain or toll-keeper} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{landslide or thin air}.
\item \textbf{Foe Appears:} A \textbf{rival raja's guard or a demon fragment} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Aeler engineers, raja's soldiers, or mountain spirits} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes—\textbf{mine seals/lease stone cracks/ninth echo spoken}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Viceroy, High Raja, or Demon from the Ninth Seal} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically—\textbf{mine collapses/bridge falls/lease stone shatters}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice—\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Satrapies (Mine Law):} Iron veins glow faintly, the mountain groans, a breath-counter ticks backward.
\item \textbf{Satrapies (Lease Politics):} The Viceroy's bell rings once for no reason, a raja's elephant kneels unprompted, a treaty scroll spontaneously combusts.
\item \textbf{Satrapies (Pass Law):} Snow falls upward, the ninth echo repeats a different name each time, a bridge appears where there was only chasm.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Thin Air}.
\item \textbf{Cut the Toll (Presence + Sway):} On a hit, freeze range for one exchange; if a toll waiver is in your possession, +1 die.
\item \textbf{Weather the Cave-in (Body + Athletics):} On a hit in \emph{Cave-In}, you pick who escapes first; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Iron \& Breath)}
\begin{itemize}
\item \textbf{The Breath-Eater (Nine Silence)} — \emph{Phases}: Mines Collapsing $\rightarrow$ Orchestrated Cave-In $\rightarrow$ Pass Coup. \emph{Cracks}: find a lost breath-counter, expose a false lease, survive a count of nine.
\item \textbf{The Unsealed (Demon of the Ninth Seal)} — \emph{Phases}: Whispers in the Mine $\rightarrow$ Procession of Cracking Timbers $\rightarrow$ Iron-Baptism of a Viceroy. \emph{Cracks}: relic Covenant seal, survivor testimony, the mine's face weeping clear water.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Mountain Satrapies, power flows through iron, tea, and the weight of the lease. The Viceroy holds authority from the Aeler High King, but the rajas control the mines, the Covenant judges the courts, and the elephant masters move the ore. No single faction rules — and every faction owes a share of their success to the mountain.

\textbf{For the GM:}  
Power in the Satrapies is measured in breath-counts, toll receipts, and the loyalty of the elephant caravans. Patronage often takes the form of iron ingots, breath-counters, or toll waivers — each tied to a specific mine, bridge, or clan. To emphasise community and leverage over debt:
\begin{itemize}
\item Tie rewards to physical tokens (bells, ingots, nails) that can be challenged, stolen, or voided by a rival.
\item Let the lease negotiation be a campaign front: the party's actions tick clocks for Aeler, rajas, Admiralty, and Covenant.
\item Use the mines, bridges, and the Viceroy's hall as arenas for social contests, where a breath-count is a threat and a sealed mine is a death sentence.
\item Encourage players to earn \textbf{Share Tokens} from the elephant caravans — these are not debts but markers of mutual trust.
\end{itemize}
In the Mountain Satrapies, the mountain forgets nothing — and the ninth breath is always waiting. But the mountain also remembers every kindness, every shared fire, every elephant that was fed.
\end{tcolorbox}

% Index entries
\index{Mountain Satrapies|see{Aeler-Dhaharan}}
\index{Theme generator}
\index{Aeler-Dhaharan generator}
\index{Places!Satrapies}
\index{People!Satrapies}
\index{Complications!Satrapies}
\index{Rewards!Satrapies}
\index{Viceroy Thrain}
\index{Rani Jaya}
\index{Sister Gaura}
\index{Ninth seal mine}
\index{Elephant caravans}
\index{Iron passes}
\index{Breath-counter}
\index{Lease stone}
\index{Local Patrons!Satrapies}