\section{Ngomebe --- ``The Stone Men''}
\label{chap:ngomebe}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The stone remembers. The ancestors walk with the rollers. The city that stops dies.}

\begin{quote}
  \textbf{Mason-Kin Elder (Elite):}  
  ``A city that stops moving is a city that has chosen to die. The stone remembers every mile, every roller, every oath sworn in the shadow of the keystone. We do not build cities. We raise them from the earth and teach them to walk.''
\end{quote}

\begin{quote}
  \textbf{Stone-Cutter (Commoner):}  
  ``My grandfather carved the keystone for this city. My father greased the rollers. I inherited a sledge and a song. The song is the city's heartbeat. If I stop singing, the city stops walking --- and so do I.''
\end{quote}

\begin{quote}
  \textbf{Ancestral Healer (Izinyanga):}  
  ``The ancestors do not speak in words. They speak in cracks. A fissure in the keystone is not a flaw --- it is a message. Learn to read the stone's wounds, or the city will read your bones.''
\end{quote}

\textbf{Genre / Mood:} Walking city survival, ancestral stone magic, industrial partnership, ecological heresy.

\textbf{Starting Location:} The edge of a moving city's path, where the stone is still warm and the dust has not yet settled. A Mason-Kin elder approaches, wiping granite dust from his face: ``A keystone has cracked. The city will stop before dawn. The ancestors are restless. Find the saboteur --- a Root-binder, a tokoloshe, or a traitor among us --- or push the rollers yourselves. The plateau does not forgive stillness.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Ngomebe are a people of deep, iron-dark skin and older ties to the southern ranges. Their cities walk --- colossal stone structures on rollers and sledges, dragged by teams of oxen, magic, and Aeler-designed capstans. This partnership is one of the great cross-cultural achievements of the age: Aeler engineers maintain the bearings; Ngomebe masons carve the keystones. A walking city is not a curiosity; it is a living argument against stillness, a monument to survival in a land where staying in one place invites famine, war, or the slow forgetting of ancestral stone.

Spirits are not distant --- they live in the rock, the rollers, the cracks. The ancestors speak through keystone fissures. The \textbf{tokoloshe} fouls the bearings at night. The \textbf{impundulu} (lightning bird) shatters entire sledges with a single strike. The Ngomebe do not fight these spirits. They bargain with them: a grindstone of fresh meat, a pot of home-brewed sorghum beer, a promise whispered into a roller-pit.

The Ngomebe are no more united than the stone is silent. \textbf{Mason-Kin Houses} feud openly over migration routes and carving rights. The \textbf{Aeler Covenant} guards its technical secrets even as it swears partnership. \textbf{Root-binders} gather in the groves, preaching that motion is a curse. Among the \textbf{Sledge-Born} laborers, songs of stillness spread in whispers. And the \textbf{Izinyanga} keep their own counsel, reading cracks that houses and engineers would rather ignore.

\vspace{2mm}
\textbf{Neighbors and Rivals:}
\begin{itemize}
  \item \textit{Northwest --- Sekogo:} The Ukwe's roots reach into Ngomebe's southern savanna. Sekogo druids trade honey for keystone shards, believing the stone contains trapped memories of the jungle's oldest trees.
  \item \textit{Northeast --- Aelinnel of the Southern Coast:} Fae-touched gnomes who broker the pact that keeps the walking cities from sinking into soft ground. Their dawn bone-roll determines each day's safe path. The Ngomebe tolerate them; the gnomes hold a bond older than memory.
  \item \textit{East/Southeast --- Aelaerem of Taharka (Highland halflings):} Small, red-threaded folk who tend the passes and offer shelter to weary caravans. They share a reverence for ancestral stone and trade coffee for iron.
  \item \textit{South --- The Salt Scar:} A vast, cracked waste where abandoned keystones lie like frozen waves. The \textbf{Still Voice} haunts these ruins, asking every traveler why their city stopped.
\end{itemize}

\vspace{2mm}
\textbf{What you notice first:} The groan of stone on stone. The smell of warm granite, ox sweat, and sorghum beer offerings. The way the ground trembles slightly, even a mile away, as the city inches across the plateau.

\textbf{The one rule that kills newcomers:} Never carve a name into living rock. The stone remembers. It will whisper that name to every spirit of the earth, and the weight will follow your bloodline for seven generations.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of Movement}

Power here is measured in keystones, migration routes, and the loyalty of the Mason-Kin. A walking city is a polity, a temple, and a weapon all at once. Yet its people are far from united. The \textbf{Mason-Kin Houses} control the rights to carve, to move, and to settle, but they feud endlessly over precedence and pride. The \textbf{Root-binders}---heretics who plant acorns in the rollers---preach that the cities should stop and let the forest reclaim the stone, and their numbers swell among the weary. The \textbf{Aeler Covenant} engineers are neither masters nor servants; they are partners in a bond older than memory, though some among them whisper that the Ngomebe would be lost without Aeler steel. Beneath them all, the \textbf{Izinyanga} speak for the dead, and the dead speak through the keystone's cracks. Even the \textbf{Sledge-Born}---the ox-drivers and grease-carriers---have begun to murmur that the weight of motion is too heavy to bear forever. When a city stops, it is because the ancestors have been offended. When a city moves true, it is because the \textbf{Rain Queen} has blessed the path.

\subsection*{The Walking City Timers (Campaign Engines)}
Every city that moves is a living system. Track its health with these visible timers:

\textbf{Migration Timer [10]} --- Advances each day the city moves, retreats when it stops. At 2: fertile land $\rightarrow$ +1 die to foraging. At 6: narrow pass $\rightarrow$ navigation Desperate. At 10: destination reached $\rightarrow$ Keystone Blessing.  
\textbf{Roller Integrity [8]} --- Ticks when rollers crack or are sabotaged. At 4: speed halves. At 8: the city stops.  
\textbf{Ancestor Favor [8]} --- Starts at 4. Ticks up when the ancestors are honored, down when they are offended. At 0: the Still Voice comes. At 8: the ancestors bless the city (one free Keystone Sense use per scene).  
\textbf{Root-binder Attention [6]} --- Ticks when the party interacts with Root-binders or damages the forest. At 6: the Root-binders attack.

\subsection*{The Still Voice ($1\text{d}6$)}
When Ancestor Favor reaches 0 or the city stops, the Still Voice appears. Roll $1\text{d}6$:
\begin{enumerate}
\item ``Why did you stop?'' Answer truthfully, or it takes a memory (lose 1 XP, redeemable by a quest).
\item The Voice points at a character. That character now carries a \textbf{Dead Keystone Shard} --- it weighs nothing, but the Hollow knows where they are.
\item The ground beneath the city cracks. The city sinks one foot. Roller Integrity $-2$.
\item The Voice demands a story about a city that never stopped. If you cannot tell one, it takes the name of your home village.
\item A keystone fragment rises from the earth. Touch it to ask one question about the city's past. The Voice answers in a whisper.
\item The Voice weeps dust. The dust is bitter. If you taste it, you see the apocalypse that made the cities walk. Mark 1 Corruption.
\end{enumerate}

\subsection*{The Cracking Wind (Umbhemisi)}
The Salt Scar is not a desert. It is the graveyard of cities that stopped walking. Now a red dust --- not sand, but ground bone and forgotten oaths --- has begun to crawl north. The elders call it \textbf{Umbhemisi}, the Cracking Wind. Where it touches a keystone, the stone weeps rust. Where it touches a roller, the wood splits. The cities that hear its song are moving south, but the dust follows.

\textbf{Umbhemisi's Advance [8]} --- A campaign timer that ticks when:
\begin{itemize}
\item The party remains in one place for more than a week (+1)
\item A keystone is cracked by sabotage or neglect (+1)
\item The Still Voice is summoned (+2)
\item A city stops moving (+2)
\end{itemize}
At 4 segments: Red dust storms appear on the horizon. All navigation rolls suffer $-1$ die.  
At 8 segments: Umbhemisi reaches the city. Roller Integrity ticks $-1$ per day, and the Cracking Wind itself may manifest as a boss (use the \textbf{Dust Walker} stats from the Nine Rollers sidebox).

\subsection*{Prominent NPCs of Ngomebe}
\label{sec:ngomebe-npcs}

The walking cities host several competing factions. The \textbf{Mason-Kin Houses} guard tradition and migration rights. The \textbf{Aeler Covenant} maintains the mechanical bond. The \textbf{Izinyanga} speak for the ancestors. The \textbf{Root-binders} preach stillness. The \textbf{Sledge-Born} are the common laborers who grease the rollers and often question the cost of endless motion.

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Elder Morvath & Mason-Kin elder of the eastern caravan & Keep the city moving south, away from the Cracking Wind & He has not slept in forty days; the city's rhythm keeps him awake. \\
\hline
Taylor & Stone-cutter whose city left him behind & Catch up to his family's city before the Salt Scar claims it & He chisels a new keystone every night. None have been accepted. \\
\hline
Vent-Prior Thrain & Aeler engineer, grease-stained and level-carrying & Maintain the bearings that keep the city alive & She speaks to the rollers in a language of taps. They answer in groans. \\
\hline
The Root-binder & A heretic planting acorns in the roller-pits & Let the forest reclaim the stone; let the cities sleep & He has a single green leaf growing from his left ear. \\
\hline
Ox-Master Senn & Handler of the eight-ox teams that drag the city & Keep the beasts healthy, the yokes oiled, and the pace steady & He names each ox after a dead mason. The oxen remember. \\
\hline
The Still Voice & A spirit of abandoned keystones & Witness the cities that chose to stop & It appears only when a city's migration timer reaches zero. It asks one question: ``Why did you stay?'' \\
\hline
Nomvula (Rain Queen) & Mythical rainmaker, keeper of the drought & Ensure the plateau never dries while the cities walk & She has not been seen in a century. Her staff is a petrified lightning bolt. \\
\hline
Tokoloshe Catcher & Old woman who traps water spirits & Keep the bearings free of malevolent tricksters & She catches tokoloshe in clay pots; they sing when the pot is shaken. \\
\hline
\end{longtable}

\subsection*{Names of Ngomebe}
\label{sec:ngomebe-names}

Inspired by the deep southern plateau: short, earth-toned syllables with clicks and tonal shifts. Surnames often denote a stone type, a city of origin, or a Mason-Kin House.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Morvath & Thandi & Keystone-carver & \textit{the Unstill} \\
\hline
Taylor & Nia & Roller-born & \textit{the Gray-Faced} \\
\hline
Senn & Zola & Mason-Kin & \textit{the Slow Walker} \\
\hline
Duma & Amara & Stone-song & \textit{the Ninth Roller} \\
\hline
Temba & Lindi & Sledge-heir & \textit{the Root-Breaker} \\
\hline
Jengo & Penda & Ox-kin & \textit{the Warm Stone} \\
\hline
Nomvula & Lethabo & Rain-touched & \textit{the Drought's Daughter} \\
\hline
\end{longtable}

\paragraph*{(Quarry/Hall/Edge)} Abandoned quarry where stone-song still echoes; Mason-Kin hall with walls carved in lineage; the edge of a city's path --- the stone is warm to the touch.

\section*{Spades --- Places}
\label{sec:ngomebe-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{City on the Move} --- Rollers groan, oxen low, the ground trembles.
  \hfill \textit{The city's shadow moves faster than the city itself. Stay in the shadow to keep pace.} \\
  \textbf{Mechanical hook:} If you fall out of the shadow, tick \textbf{Migration Timer $-1$} (the city leaves you behind). On a Miss, you are separated for one scene.

\item \textbf{Abandoned Quarry} --- Stone-song echoes from the walls.
  \hfill \textit{The song is the memory of the last stone cut here. It sings of a mason who fell.} \\
  \textbf{Mechanical hook:} A character who listens (Spirit+Lore DV 4) learns a secret of the stone (a Clue about a forgotten city). On failure, the song haunts them --- gain the Condition \textit{Quarry Dreams} ($-1$ die on all Resolve tests for a day).

\item \textbf{Mason-Kin Hall} --- Walls carved with lineage, floor inlaid with sledge-ruts.
  \hfill \textit{The carvings weep water when a city is in danger. The water tastes of granite.} \\
  \textbf{Mechanical hook:} Drinking the water restores 1 Fatigue but ticks \textbf{Ancestor Favor $-1$} (the ancestors are watching). Refusing to drink when offered ticks \textbf{Ancestor Favor $-2$}.

\item \textbf{Edge of the Path} --- The stone is warm. The warmth fades when the city leaves.
  \hfill \textit{If you place your palm on the stone before it cools, you may ask the city one question.} \\
  \textbf{Mechanical hook:} Ask the GM one yes/no question about the city's destination. On a Miss, the stone burns your hand --- mark 1 Fatigue and the answer is a lie.

\item \textbf{Roller-Pit} --- A trench where the great rollers turn, greased with rendered ox-fat.
  \hfill \textit{The grease is mixed with ancestor ash. The rollers remember every mason who ever touched them.} \\
  \textbf{Mechanical hook:} Entering the pit requires Body+Athletics DV 4. On failure, take 1 Harm. Sabotaging the pit (Wits+Craft DV 5) ticks \textbf{Migration Timer $-2$} but starts a \textbf{Root-binder Sabotage [4]} timer.

\item \textbf{Keystone Chamber} --- The heart of the city, a single stone that bears all weight.
  \hfill \textit{The keystone hums. The pitch changes when a mason swears a false oath.} \\
  \textbf{Mechanical hook:} A character who swears an oath here cannot break it without ticking \textbf{Ancestor Favor $-3$}. The keystone will also reveal one lie spoken in its presence per scene.

\item \textbf{Sledge-Track Scar} --- A groove worn into the bedrock by centuries of passage.
  \hfill \textit{The groove is deep enough to hide a corpse. Some say it already does.} \\
  \textbf{Mechanical hook:} A character who hides in the scar gains +2 dice to Stealth, but on a Miss they disturb a ghost --- start a \textbf{Haunted Track [4]} timer.

\item \textbf{Oxen Corral (Mobile)} --- A pen on sledges, dragged behind the city.
  \hfill \textit{The oxen never sleep. They walk in their sleep. Their dreams move the city.} \\
  \textbf{Mechanical hook:} A character who soothes the oxen (Presence+Beastcraft DV 4) clears 1 Fatigue from the team and advances \textbf{Migration Timer +1}. On failure, a stampede --- lose 1 Supply.

\item \textbf{Root-binder's Grove} --- A hidden stand of trees planted in an abandoned roller-pit.
  \hfill \textit{The trees have stone roots. The roots grip the rollers and will not let go.} \\
  \textbf{Mechanical hook:} Entering the grove ticks \textbf{Root-binder Attention +1}. A character who offers a story (Spirit+Performance DV 4) may gain a \textbf{Green Blessing} (one use of acorn magic). On failure, the roots grab them --- test Body+Athletics or be pinned.

\item[J] \textbf{Abandoned City (Dead Keystone)} --- A stone shell on a hill, silent and cold.
  \hfill \textit{The keystone is cracked. A single green shoot grows from the crack. The Root-binder's work.} \\
  \textbf{Mechanical hook:} The Still Voice is here. Roll $1\text{d}6$ on the Still Voice table. If the party takes the shoot, they may plant it to start a new grove --- but the Hollow will know.

\item[Q] \textbf{Bearing Foundry (Aeler)} --- A mobile forge where new rollers are cast.
  \hfill \textit{The forge is pulled by a separate team of oxen. It is always last in the procession.} \\
  \textbf{Mechanical hook:} The foundry can repair one roller per day (clear 2 segments of Roller Integrity) but costs 1 Supply in fuel. If the party is unwelcome, the Aeler engineer demands a story or a favor.

\item[K] \textbf{The Thing-Holm of Stone} --- A flat rock where Mason-Kin elders meet.
  \hfill \textit{The rock is a single piece of fallen sky-metal. Oaths sworn here cannot be broken.} \\
  \textbf{Mechanical hook:} Any oath sworn here is witnessed by the ancestors. Breaking it ticks \textbf{Ancestor Favor $-4$} and the oath-breaker is exiled from all walking cities.

\item[A] \textbf{The First City (Myth)} --- A walking city that never stopped. It is said to circle the plateau, just beyond the horizon.
  \hfill \textit{Those who see it cannot speak of it for a year and a day. The stone seals their lips.} \\
  \textbf{Mechanical hook:} If the party sights the First City, they gain a \textbf{First City's Echo} (Diamond). The GM also gains 3 SB --- the Hollow is watching through the city's eyes.
\end{enumerate}

\paragraph*{(Cutter/Elder/Engineer)} Stone-cutter whose city left them behind; Mason-Kin elder, gray with granite dust; Aeler engineer: grease-stained, carrying a level.

\section*{Hearts --- People \& Factions}
\label{sec:ngomebe-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Stone-Cutter (Left Behind)} --- Her city moved south. She stayed to carve a monument.
  \hfill \textit{She carves her own face into the rock. ``So the city will remember me,'' she says.} \\
  \textbf{Obligation:} She carries a sledge-right to a House that no longer exists. If you find the House's charter, she will teach you a stone-song. \\
  \textbf{Cost of aid:} A chisel strike on her monument. She will not say where.

\item \textbf{Mason-Kin Elder} --- Gray with granite dust, eyes the color of slate.
  \hfill \textit{He carries a hammer that has never struck a false blow. The hammer knows.} \\
  \textbf{Obligation:} He lost a keystone charter to a Root-binder. If you recover it, he will grant you a \textbf{Keystone Charter}. \\
  \textbf{Cost of aid:} A day's labor in the roller-pits (Body+Athletics DV 4). On failure, you owe a story about a city that fell.

\item \textbf{Root-binder Heretic} --- Planting acorns in the roller-pits, singing to the earth.
  \hfill \textit{His hands are bark. His breath smells of green wood. The oxen fear him.} \\
  \textbf{Obligation:} He was once a mason. The ancestors cursed him. If you lift the curse (a ritual at the Keystone Chamber), he will give you a \textbf{Green Pact}. \\
  \textbf{Cost of aid:} An acorn from his grove. Planting it ticks \textbf{Root-binder Attention +2}.

\item \textbf{Aeler Engineer} --- Tuning fork in one hand, level in the other.
  \hfill \textit{She can hear a crack in a bearing from a mile away. She never walks under a keystone.} \\
  \textbf{Obligation:} She lost her tuning fork in a roller-pit. Find it, and she will repair any bearing for free. \\
  \textbf{Cost of aid:} A story about a machine that failed. She is collecting evidence for a case against the High Council.

\item \textbf{Ox-Master} --- Whip of braided leather, voice like a landslide.
  \hfill \textit{He names his oxen after dead masons. The oxen stop when he speaks their names.} \\
  \textbf{Obligation:} One ox is sick. He needs a herb that grows only in Root-binder groves. If you fetch it, he will give you an \textbf{Ox-Master's Whip}. \\
  \textbf{Cost of aid:} A day of driving the oxen. On a Miss, the team bolts --- tick \textbf{Migration Timer $-1$}.

\item \textbf{Ancestral Healer (Izinyanga)} --- Reads the cracks in the keystone, speaks for the dead.
  \hfill \textit{She wears a necklace of ground quartz. The ancestors whisper through the stones.} \\
  \textbf{Obligation:} The ancestors have gone silent. She needs a sacrifice --- a memory of a city that fell. If you give one, she will interpret a keystone crack for you (gain a Clue). \\
  \textbf{Cost of aid:} A drop of your blood on the keystone. The ancestors will remember your face.

\item \textbf{Tokoloshe Catcher} --- Old woman, blind in one eye, carries a clay pot.
  \hfill \textit{She traps tokoloshe in her pots. They sing when shaken. She uses their songs to find sabotage.} \\
  \textbf{Obligation:} A tokoloshe escaped. Catch it, and she will give you a \textbf{Tokoloshe Pot} (once, +1 die to detect sabotage). \\
  \textbf{Cost of aid:} A day of listening. The tokoloshe's song will give you nightmares (GM gains 1 SB).

\item \textbf{Impundulu Tamer} --- A legendary bird-whisperer, with scars from lightning.
  \hfill \textit{He can call the lightning bird with a whistle. It demands payment: the blood of a mason.} \\
  \textbf{Obligation:} He is bound to the bird. If you pay its price (1 Supply of raw meat), he will call the bird for you once. \\
  \textbf{Cost of aid:} A scar. The bird will remember your face --- the GM may spend 1 SB to have it appear in a storm.

\item \textbf{Impi Captain} --- Warrior of the city's flank guard, carrying a shield of braided leather.
  \hfill \textit{She lost two fingers to a Root-binder ambush. She still holds her spear with the remaining three.} \\
  \textbf{Obligation:} She wants revenge on the Root-binder who took her fingers. If you help her, she will fight beside you in any city battle. \\
  \textbf{Cost of aid:} A spear throw. On a Miss, you hit an ox --- tick \textbf{Roller Integrity $-1$}.

\item[J] \textbf{The First Mason's Ghost} --- A translucent figure in a mason's apron.
  \hfill \textit{He appears when a keystone is first laid. He offers advice. The advice is always wrong.} \\
  \textbf{Obligation:} He wants his name carved on a keystone. If you do, he will grant you a \textbf{Stone-Song Fragment}. \\
  \textbf{Cost of aid:} His advice. Follow it, and the keystone will crack (tick \textbf{Roller Integrity $-2$}).

\item[Q] \textbf{The Ninth Ox} --- A mythic beast said to pull the First City.
  \hfill \textit{Its hooves never touch the ground. It walks on the shadow of clouds.} \\
  \textbf{Obligation:} It has no obligation --- it is a legend. If you see it, you may ask one question about the First City. The answer is always a riddle.

\item[K] \textbf{Bearing-Caster (Aeler)} --- A dwarf who pours molten metal into roller-molds.
  \hfill \textit{He signs each bearing with his own thumbprint. The thumbprint is a prayer.} \\
  \textbf{Obligation:} He lost his mold to a tokoloshe. Recover it, and he will cast a bearing for free. \\
  \textbf{Cost of aid:} A thumbprint. He will remember your face --- the GM may spend 1 SB to have his prayer answered against you.

\item[A] \textbf{The Still Voice} --- A spirit made of silence and cold stone.
  \hfill \textit{It asks: ``Why did you stop?'' Answer truthfully, or it will never let you leave.} \\
  \textbf{Obligation:} It collects the names of cities that stopped. If you give it a name, it will answer one question about that city's fate. \\
  \textbf{Cost of aid:} A name. The Still Voice will whisper it to the Hollow --- the GM gains a String on the city you named.
\end{enumerate}

\paragraph*{(Crack/Sabotage/Tremor)} A keystone cracks --- the city stops; Root-binder sabotage --- a wheel jams; an earth tremor (all actions Desperate for a scene).

\section*{Clubs --- Complications/Threats}
\label{sec:ngomebe-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Cracked Keystone} --- The heart of the city fractures. The city stops.
  \hfill \textit{The crack is singing. The song is the name of the mason who carved it. He is long dead.} \\
  \textbf{Mechanical hook:} Tick \textbf{Roller Integrity $-2$}. If Roller Integrity reaches 0, the city stops and the Still Voice appears (roll on its table). The crack can be sealed with a mason's promise (a story or a memory).

\item \textbf{Root-binder Sabotage} --- Acorns sprout in the roller-pits. A wheel jams.
  \hfill \textit{The roots are intelligent. They twist toward the sound of oaths.} \\
  \textbf{Mechanical hook:} Tick \textbf{Root-binder Attention +1}. To clear the jam, a character must swear an oath (any oath). Swearing ticks \textbf{Ancestor Favor $-1$}. Breaking it later ticks $-3$.

\item \textbf{Earth Tremor} --- The ground shakes. All actions are Desperate for a scene.
  \hfill \textit{The tremor was not natural. The Still Voice is waking.} \\
  \textbf{Mechanical hook:} Tick \textbf{Ancestor Favor $-1$}. On a Miss, a roller slips --- tick \textbf{Roller Integrity $-1$}. The party may offer a sacrifice (1 Supply) to calm the tremor.

\item \textbf{Mason-Kin Feud} --- Two Houses argue over the next migration route.
  \hfill \textit{The argument is carved into the keystone. The stone is splitting.} \\
  \textbf{Mechanical hook:} The party must choose a side or mediate. Choosing a side gives an ally but ticks \textbf{Migration Timer $-2$} (the other House sabotages the route). Mediating requires Presence+Sway DV 5; on failure, both Houses turn against you.

\item \textbf{Oxen Plague} --- A sickness spreads through the teams. Speed halves.
  \hfill \textit{The sickness came from a Root-binder's grove. The leaves are in the fodder.} \\
  \textbf{Mechanical hook:} Tick \textbf{Migration Timer $-1$} per day until cured. Cure requires a Medicine test (DV 5) or a Root-binder's blessing (costs 1 Supply and ticks \textbf{Root-binder Attention +1}).

\item \textbf{Bearing Failure (Aeler)} --- A roller seizes. The engineer blames the stone-cutter.
  \hfill \textit{The bearing was sabotaged with dust from the Salt Scar. The dust whispers.} \\
  \textbf{Mechanical hook:} Tick \textbf{Roller Integrity $-1$}. The engineer demands a replacement bearing (1 Supply) or a public apology. Refusal starts a \textbf{Engineer's Grudge [4]} timer.

\item \textbf{Sledge-Track Collapse} --- The ground gives way. A section of the city sinks.
  \hfill \textit{The collapse reveals an older city buried beneath. The older city is still walking.} \\
  \textbf{Mechanical hook:} Tick \textbf{Migration Timer $-2$}. The party may explore the older city (a mini-dungeon). Each scene there ticks \textbf{Ancestor Favor $-1$} (the dead are restless).

\item \textbf{Migration Blockade} --- Another city has stopped on the planned route.
  \hfill \textit{The stopped city's keystone is dead. Its people have become Root-binders.} \\
  \textbf{Mechanical hook:} The party must negotiate with the Root-binders (Presence+Sway DV 5) or clear a path (Craft test, DV 6). Failure ticks \textbf{Migration Timer $-3$} (detour).

\item \textbf{Ancestral Claim Called} --- A Mason-Kin House demands a sledge-right from a century ago.
  \hfill \textit{The right is carved on a tablet of slate. The slate is warm. The ancestors are watching.} \\
  \textbf{Mechanical hook:} Pay 2 Supply or offer a hostage (a PC becomes a temporary follower of the House). Refusal ticks \textbf{Ancestor Favor $-2$} and starts a \textbf{Feud of the Slate [6]}.

\item[J] \textbf{Tokoloshe in the Bearings} --- A water spirit fouls the roller-pits. The city groans.
  \hfill \textit{The tokoloshe is small, fast, and invisible to the unaware. Only a catcher can see it.} \\
  \textbf{Mechanical hook:} Tick \textbf{Roller Integrity $-1$} per scene until caught. Catching requires a tokoloshe catcher (NPC) or a Wits+Spirit test DV 5. On failure, the tokoloshe escapes --- tick again.

\item[Q] \textbf{Impundulu Strikes} --- A lightning bird shatters a keystone with a thunderous blow.
  \hfill \textit{The bird is angry. Someone broke a vow to the Rain Queen. The city bleeds stone.} \\
  \textbf{Mechanical hook:} Tick \textbf{Roller Integrity $-3$}. The bird demands a sacrifice: the blood of a mason (1 Harm to a PC) or a vow renewal (swear an oath to the Rain Queen). Refusal ticks \textbf{Ancestor Favor $-3$}.

\item[K] \textbf{Root-binder Revolt} --- A grove of walking trees attacks the roller-pits.
  \hfill \textit{The trees have stone bark. The stones are carved with Mason-Kin faces.} \\
  \textbf{Mechanical hook:} Combat (Cap 3 adversary). If the party loses, the city stops for $1\text{d}4$ days. If they win, tick \textbf{Root-binder Attention} to zero --- the grove burns, but the Hollow's attention ticks +2.

\item[A] \textbf{The First City Approaches} --- A city older than memory appears on the horizon.
  \hfill \textit{It is made of obsidian. Its rollers are giant femurs. It wants to merge.} \\
  \textbf{Mechanical hook:} The party has one scene to flee or negotiate. Merging cities will erase both identities --- the party loses all Ngomebe Assets. If they flee, tick \textbf{Migration Timer $-5$} (the First City chases them).
\end{enumerate}

\paragraph*{(Charter/Wax/Song)} Keystone charter (right to dwell in a city for a season); sledge-wax (greases negotiation with the Mason-Kin); stone-song fragment (one reroll on a survival test).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:ngomebe-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Keystone Charter} --- Right to dwell in a city for a season, tax-free.
  \hfill \textit{The charter is a chip of the keystone itself. Carry it to be welcomed.} \\
  \textbf{Hidden cost:} If you break the charter's terms (e.g., harm a mason), the chip turns to dust and you gain the Condition \textit{Stone-Cursed} ($-1$ die on all rolls involving Ngomebe for a year).

\item \textbf{Sledge-Wax} --- A lump of gray grease. Use it to gain +1 Position in Mason-Kin negotiations.
  \hfill \textit{The wax is rendered from the fat of an ox that walked a thousand miles.} \\
  \textbf{Hidden cost:} The ox's ghost follows you. Once per session, the GM may have it low at an inopportune moment (tick \textbf{Migration Timer $-1$}).

\item \textbf{Stone-Song Fragment} --- A recorded hum. Reroll a failed Survival test once.
  \hfill \textit{The song is the city's heartbeat. Hum it, and the stone will guide you.} \\
  \textbf{Hidden cost:} Humming the song ticks \textbf{Ancestor Favor $-1$} (the dead are listening). If you use it in a city that has stopped, the Still Voice appears.

\item \textbf{Roller-Bearing (Aeler-crafted)} --- A spare bearing. Gift it to an engineer for a favor.
  \hfill \textit{The bearing is stamped with a rune that means ``no burden marked.'' The rune is a lie.} \\
  \textbf{Hidden cost:} Using the bearing voids the lie --- the engineer will remember. The GM may spend 1 SB to have the bearing fail at a critical moment.

\item \textbf{Mason-Kin Hammer} --- A tool that never misses a keystone's true face.
  \hfill \textit{The hammer's head is a piece of the First City's keystone. It whispers warnings.} \\
  \textbf{Hidden cost:} The hammer whispers to the Hollow. Each time you use it, tick \textbf{Ancestor Favor $-1$}. When Ancestor Favor reaches 0, the hammer shatters.

\item \textbf{Ox-Master's Whip} --- Crack it to make oxen move faster for one leg.
  \hfill \textit{The whip's braid includes a hair from every ox that ever pulled a city.} \\
  \textbf{Hidden cost:} The whip angers the oxen's ghosts. After use, the next stampede is automatic (GM may spend 1 SB to trigger it).

\item \textbf{Root-binder's Acorn} --- Plant it to stop a city's roller for one day.
  \hfill \textit{The acorn grows into a sapling within an hour. The sapling has stone roots.} \\
  \textbf{Hidden cost:} Planting the acorn ticks \textbf{Root-binder Attention +2} and \textbf{Ancestor Favor $-2$} (the ancestors see it as betrayal). The acorn can be removed only by a Root-binder's blessing.

\item \textbf{Abandoned City Map} --- A chart of dead cities and their keystone graves.
  \hfill \textit{The map is drawn on stone. The stone is a page. The pages are bound in granite.} \\
  \textbf{Hidden cost:} The map is haunted. Each time you consult it, roll $1\text{d}6$. On a 1, the Still Voice appears (roll on its table). On a 6, you find a forgotten keystone --- gain 1 Supply.

\item \textbf{Elder's Verdict} --- A ruling that settles a Mason-Kin dispute in your favor.
  \hfill \textit{The verdict is carved on a slate tablet. Carry it as a shield.} \\
  \textbf{Hidden cost:} The losing House will remember. Start a \textbf{Vendetta [6]} timer. When it fills, they sabotage your city.

\item[J] \textbf{Bearing-Foundry Pass} --- Access to the Aeler mobile forge for one repair.
  \hfill \textit{The pass is a single steel coin. The coin is still hot.} \\
  \textbf{Hidden cost:} The forge is a witness. If you use the pass for a dishonest repair, the coin burns your hand --- mark 1 Harm and the engineer will never trust you again.

\item[Q] \textbf{Still Voice's Silence} --- One question answered truthfully by the spirit of abandoned stone.
  \hfill \textit{The answer comes as a cold breath. The breath smells of geode crystals.} \\
  \textbf{Hidden cost:} After the answer, the Still Voice marks you. Tick \textbf{Ancestor Favor $-2$}. The Voice will ask you a question in return; you must answer truthfully or lose 1 XP.

\item[K] \textbf{First City's Echo} --- A piece of obsidian that hums with ancient rhythm.
  \hfill \textit{Once per campaign, touch it to make a city move one mile in a single night.} \\
  \textbf{Hidden cost:} Using the Echo ticks \textbf{Ancestor Favor $-4$} (the First City's ancestors are disturbed). The city that moves will be haunted by the Echo's ghost for a year (GM gains 1 SB per session).

\item[A] \textbf{The Living Keystone} --- A fragment from a city that still walks.
  \hfill \textit{Plant it in dead stone, and a new city will grow. The growth takes a century.} \\
  \textbf{Hidden cost:} The keystone is alive. It will demand a blood-price (1 Harm per year of growth). Refusal turns it to dust --- the Hollow takes your name.
\end{enumerate}

\section*{Quick use notes}
\label{sec:ngomebe-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (charters, wax, songs) that change position rather than call for a roll.
\item If any A appears, echo \textbf{stone-omens}---rollers that groan in harmony, keystones that weep water, a city that leaves no track.
\item Track the \textbf{Migration Timer [10]}, \textbf{Roller Integrity [8]}, \textbf{Ancestor Favor [8]}, and \textbf{Root-binder Attention [6]} as campaign-wide pressures.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:ngomebe-mechanics}

\begin{itemize}
\item \textbf{The Keystone Bond (Relationship Mechanic):} Each walking city has a Keystone --- a single stone that bears the weight of the roof. The Keystone is also the physical anchor of the city's ancestral spirit. A character who bonds with the Keystone (by spending a downtime scene in meditation at its base) gains:
  \begin{itemize}
    \item \textbf{Keystone Sense:} Once per scene, the character can ask the Keystone one question about the city's current state (``Is the eastern roller cracking?'' ``Is there a saboteur in the foundry?''). The Keystone answers truthfully, but only about the present.
    \item \textbf{Keystone Weight:} The character gains a 4-segment \textbf{Ancestor's Attention} timer. When it fills, the ancestor demands a service --- usually a pilgrimage to a forgotten quarry or the retrieval of a lost roller.
  \end{itemize}
\item \textbf{The Roller-Pits (Hazard Mechanic):} The city moves on rollers --- massive wooden cylinders that must be greased and rotated. The roller-pits are dangerous, loud, and prone to sabotage.
  \begin{itemize}
    \item Entering a roller-pit requires a \textbf{Body + Athletics test (DV 4)} to avoid being crushed. On a failure, the character takes 1 Harm.
    \item Sabotaging a roller-pit (to stop the city or redirect it) requires a \textbf{Wits + Craft test (DV 5)}. On a success, the city moves in the desired direction; on a failure, the roller jams and the city stops --- and the ancestors are angry.
    \item Repairing a jammed roller takes 3 actions. Each action requires a \textbf{Body + Craft test (DV 4)}. On a failure, the roller shifts and the character takes 1 Fatigue.
  \end{itemize}
\item \textbf{The Migration Timer (Campaign Timer , size 10):} Every walking city has a Migration Timer . It advances each day the city moves, and retreats when the city stops.
  \begin{itemize}
    \item At 2: The city enters fertile land. The party gains +1 die to foraging and hunting.
    \item At 4: The city passes a rival settlement. The party may trade or raid.
    \item At 6: The city enters a narrow pass. All navigation rolls are Desperate.
    \item At 8: The city approaches a sacred site. The ancestors grow restless --- all rolls involving the Keystone gain +1 die, but failure ticks the Ancestor's Attention timer by 2.
    \item At 10: The city reaches its destination --- the end of the migration. The ancestors are satisfied. The party may claim a \textbf{Keystone Blessing} (a permanent +1 die to all rolls involving stonework or negotiation with Ngomebe).
  \end{itemize}
  The party can influence the Migration Timer by:
  \begin{itemize}
    \item \textbf{Scouting ahead} (Wits + Survival) --- success advances the timer by 1, failure ticks it back by 1.
    \item \textbf{Repairing rollers} (Craft test) --- success advances the timer by 2, failure advances it by 0 but ticks the Ancestor's Attention.
    \item \textbf{Paying tribute} (a Honey Jar, a story, or a Boon) --- advances the timer by 1 without risk.
  \end{itemize}
\item \textbf{The Root-binder Heresy (Faction Mechanic):} Root-binders are heretics who plant acorns in the roller-pits, hoping to stop the cities and let the forest reclaim the stone. They have a \textbf{Sabotage Timer (size 6)}. Each time the party fails a Craft or Survival roll in a roller-pit, the Sabotage Timer ticks.
  \begin{itemize}
    \item At 2: A roller cracks. The city loses 1 Speed (travel takes longer).
    \item At 4: A roller-pit floods. The city must stop for repairs (lose 1 day).
    \item At 6: A major collapse. The city cannot move until the party clears the pit (requires a Strength + Craft test, DV 5, and costs 1 day).
  \end{itemize}
  Players can work with the Root-binders (gain information, sabotage rival cities) or against them (hunt them, protect the rollers). The Root-binders offer \textbf{Green Blessings} --- a permanent +1 die to survival in the forest, but the forest's attention tightens (the Hollow's gaze falls more often).
\item \textbf{The Mason-Kin Houses:} The walking cities are not nations but extended families. Each House controls a keystone, a migration route, and the rights to carve. Houses feud over sledge-tracks, water rights, and the honor of the First City.
\item \textbf{The Aeler Pact:} The Aeler engineers of the southern ranges maintain the bearings and capstans that make the walking cities possible. The Ngomebe provide keystone hearts and stone-song. It is a partnership of equals, sealed in slate and iron.
\item \textbf{Ancestral Cracks:} Once per session, a character may press their ear to a keystone and roll Spirit + Lore (DV 3). On success, the ancestors whisper a warning or a clue (GM reveals a hidden complication or a shortcut). On a miss, the stone is silent --- but the ancestors are listening.
\item \textbf{Tokoloshe Pots:} A character who carries a clay pot with a tokoloshe inside gains +1 die to detect sabotage or traps in the city. The tokoloshe sings when danger is near. However, if the pot breaks, the tokoloshe escapes and will haunt the carrier for a full season (all actions Desperate at night).
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The First Stone} --- The Sealed Gate. \textit{The keystone that never cracks; the oath that never breaks.}
    \item \textbf{The Rain Queen} --- Ra\'eyn. \textit{The bringer of storms and the drought; the mistress of the path.}
    \item \textbf{The Lightning Bird} --- The Storm King. \textit{The shatterer of false keystones; the judge of masons.}
    \item \textbf{The Ancestor's Voice} --- The Witness. \textit{The whisper in the crack; the memory in the stone.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{Umbhemisi and the Salt Scar:} The Cracking Wind is said to be the ghost of the First City, ground into dust by its own stubbornness. Some Ngomebe elders believe that if a city ever stops moving for nine generations, the Wind will reclaim it entirely. The Root-binders whisper that the Wind is not a curse but a teacher: it cracks only the stone that refuses to grow.
  \item \textbf{The Sledge-Wax Trade (Way of Silk):} Sledge-wax is a valuable commodity on the caravan routes. Tulkani merchants trade it for story-knots; Fhara water-clerks use it to seal well-contracts. A missing shipment of sledge-wax could stop a city.
  \item \textbf{The Iron Mines of Vajrapur (Dhahara):} The Aeler engineers of Ngomebe source their bearing-metal from the Dhaharan iron mines. A dispute between a Mason-Kin House and a Dhaharan raja could choke the supply of rollers.
  \item \textbf{The Hollow and the Still City (Eastern Realms):} The Hollow is attracted to still places. A city that stops moving becomes a beacon. The Aelaerem whisper that the First City never stopped because it is running from something --- and that something may be the Cracking Wind itself.
  \item \textbf{Tokoloshe and the Bitter Root:} Daughters of the Cord sometimes capture tokoloshe and weave them into Hollowed children. The tokoloshe inside the child sings when a weight is overdue. The Ngomebe avoid Daughter-run hospices for this reason.
  \item \textbf{Sekogo's Ukwe Roots:} The Ukwe's roots are said to extend beneath the southern savanna. A Ngomebe mason who dreams of roots cracking her keystone knows that a bond to the Speaking Tree has come due.
  \item \textbf{Aelinnel Bone-Spinners:} The coastal gnomes who roll the dawn bones never explain why a failed roll sometimes sounds like a cracked keystone. The Ngomebe have learned to listen to both.
\end{itemize}

\subsection*{Ngomebe Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Roller:} Crumble a crust into the grease-pit. Ignore the first \emph{Earth Tremor} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Sledge-Rope:} Tie a single knot in an ox's halter. Cancel \emph{Oxen Plague} or take $-1$ die on your next handling roll (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Keystone:} Whisper a dead mason's name into the keystone's crack. Gain +1 Effect when negotiating with the Mason-Kin this scene, but the ancestors will expect an answer. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:ngomebe-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Mason-Kin guards or Aeler engineers}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{ox-drivers or stone-cutters}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{Root-binders or keystone guardians}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable migration pace}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local Mason-Kin elder or Aeler vent-prior} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{earth tremor or roller jam}.
\item \textbf{Foe Appears:} A \textbf{Root-binder cell or rival House guard} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Mason-Kin, Root-binders, or Aeler engineers} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes---\textbf{keystone cracks/city stops/First City sighted}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Mason-Kin High Council, Still Voice, or First City's herald} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically---\textbf{tremor opens a chasm/city splits in two/keystone begins to sing}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice---\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Ngomebe (Stone Law):} Keystone hums a warning, roller-grooves glow faintly, the ground trembles without quake.
\item \textbf{Ngomebe (Root-binder Heresy):} Acorns sprout from cracks, vines write warnings in the dust, trees move against the wind.
\item \textbf{Ngomebe (Aeler Pact):} Bearings click in counterpoint, tuning forks ring without being struck, the smell of hot metal hangs in still air.
\item \textbf{Ngomebe (Ancestral Spirits):} A cold breath whispers a name, the keystone weeps water, a tokoloshe laugh echoes from a dry roller-pit.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{False Trail}.
\item \textbf{Cut the Sledge (Presence + Sway):} On a hit, freeze range for one exchange; if a keystone charter is in your possession, +1 die.
\item \textbf{Weather the Tremor (Body + Athletics):} On a hit in \emph{Earth Tremor}, you pick who stays standing; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Stone \& Root)}
\begin{itemize}
\item \textbf{The Root-King (Nine Groves)} --- \emph{Phases}: Acorn Sabotage $\rightarrow$ Orchestrated Stillness $\rightarrow$ Forest Coup. \emph{Cracks}: burn a grove, expose a false acorn, survive a landslide.
\item \textbf{The Still Voice (Silence of Stone)} --- \emph{Phases}: Abandoned Keystones $\rightarrow$ Procession of Dead Cities $\rightarrow$ Stillness Baptism of an Elder. \emph{Cracks}: relic stone-song, survivor testimony, the city moving against the Voice.
\item \textbf{The Tokoloshe King (Master of Tricksters)} --- \emph{Phases}: Bearings Foul $\rightarrow$ Roller Pit Flooded $\rightarrow$ Oxen Stampede. \emph{Cracks}: trap him in a clay pot, negotiate with a pot of sorghum beer, feed him a false name.
\item \textbf{Umbhemisi (The Cracking Wind)} --- \emph{Phases}: Dust on the Horizon $\rightarrow$ The Song of Rust $\rightarrow$ The Stillness of Bone. \emph{Cracks}: offer a forgotten keystone, speak the name of a city that never stopped, or let a mason carve a new oath into the Salt Scar.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Ngomebe, power flows through keystones, migration rights, and the loyalty of the Mason-Kin Houses. An elder's authority rests on the age of his House's keystone and the number of cities that follow his route. The Aeler engineers hold technical authority; the Root-binders offer a heretical alternative; and the Ancestral Healers speak for the dead, who still vote through the cracks in the stone.

\textbf{For the GM:}  
Power in Ngomebe society is measured in movement and stone. Patronage often takes the form of keystone charters, sledge-wax, or stone-song fragments---each tied to a specific city and House. To emphasize this:
\begin{itemize}
\item Tie rewards to physical tokens (chips of keystone, lumps of wax, recorded songs) that can be challenged, stolen, or voided.
\item Let rival Houses issue conflicting claims, forcing players to choose whose city shelters them.
\item Use the roller-pits, keystone chambers, and Mason-Kin halls as arenas for social contests, where migration routes are debated and saboteurs are judged.
\item Introduce \textbf{Ancestral Cracks} as a way for players to gain clues from the dead --- but at the cost of the ancestors' attention.
\end{itemize}
In Ngomebe society, the stone remembers everything --- including the weight you carry. The ancestors do not forget. The tokoloshe does not forgive. And the city that stops dies.
\end{tcolorbox}

\clearpage
\begin{tcolorbox}[colback=black!3,colframe=green!60!black,title={The Nine Rollers: A City on the Edge}]

  \subsection*{The Hook}
  
  The party arrives at the edge of a moving city --- not one of the great metropolises, but a smaller Walker named \textbf{Khomani's Progress}, named after the mason who carved its keystone three centuries ago. The city is dying.
  
  \textit{Elder Morvath} meets you at the sledge-track, his face gray with granite dust and exhaustion. ``The city has not stopped in forty days. We cannot stop. The ancestors will not let us. But the rollers are cracking. The oxen are dying. And something is following us --- something that walks in the dust and counts our breaths.''
  
  He leads you to the keystone chamber. The great stone is warm. Too warm. A network of fine cracks spreads from its center like veins. The cracks are singing --- a low, mournful hum in a language none of you recognize.
  
  ``The Ninth Roller is failing,'' Morvath says. ``It was carved by Khomani herself. It has never been replaced. If it fails, the city stops. If the city stops, the Hollow comes. We have nine days until the roller cannot turn. Find a replacement --- or learn to sing the stone-song well enough to heal it.''
  
  \vspace{2mm}
  \subsection*{The Mechanics}
  
  The party has \textbf{nine days} (or scenes) to save \textit{Khomani's Progress}. Each day, they may choose one action from the table below. The city's condition is tracked by three timers:
  
  \begin{longtable}{|p{0.3\textwidth}|p{0.5\textwidth}|}
  \hline
  \textbf{Clock} & \textbf{Description} \\
  \hline
  \textbf{Roller Integrity [9]} & Starts at 9 (the Ninth Roller). Ticks down each day. At 0, the roller fails. \\
  \hline
  \textbf{Ancestor Favor [8]} & Starts at 4. Ticks up when the party honors the ancestors, ticks down when they offend. At 8, the ancestors help. At 0, the Still Voice comes. \\
  \hline
  \textbf{Root-binder Attention [6]} & Ticks up when the party interacts with Root-binders or damages the forest. At 6, the Root-binders attack. \\
  \hline
  \end{longtable}
  
  \vspace{2mm}
  \subsection*{Daily Actions}
  
  Each day, the party chooses \textbf{one} primary action and may assign up to \textbf{two} party members to support (each support grants +1 die to the primary roll).
  
  \begin{longtable}{|p{0.3\textwidth}|p{0.25\textwidth}|p{0.35\textwidth}|}
  \hline
  \textbf{Action} & \textbf{Roll} & \textbf{Outcome} \\
  \hline
  \textbf{Search for a new roller} & Wits + Survival (DV 5) & \textit{Success}: Find a suitable stone. \textit{Partial}: Find one that needs carving. \textit{Fail}: Find nothing --- tick Roller Integrity $-1$. \\
  \hline
  \textbf{Carve a replacement} & Craft + Spirit (DV 6) & \textit{Success}: New roller ready in 2 days. \textit{Partial}: Ready in 4 days. \textit{Fail}: Crack the stone --- Roller Integrity $-2$. \\
  \hline
  \textbf{Negotiate with Root-binders} & Presence + Sway (DV 4) & \textit{Success}: Root-binder Attention $-2$. \textit{Partial}: $-1$. \textit{Fail}: Root-binder Attention +2. \\
  \hline
  \textbf{Appease the ancestors} & Spirit + Lore (DV 4) & \textit{Success}: Ancestor Favor +2. \textit{Partial}: +1. \textit{Fail}: $-1$. \\
  \hline
  \textbf{Repair existing rollers} & Craft + Body (DV 4) & \textit{Success}: Roller Integrity +1. \textit{Partial}: No change. \textit{Fail}: $-1$. \\
  \hline
  \textbf{Scout the pursuing threat} & Wits + Stealth (DV 5) & \textit{Success}: Learn what follows (reveals a weakness). \textit{Partial}: Learn its nature but not its weakness. \textit{Fail}: It notices you --- next day's actions at $-1$ die. \\
  \hline
  \textbf{Study the stone-song} & Spirit + Lore (DV 6) & \textit{Success}: Learn to hum the healing melody (can restore Roller Integrity without carving). \textit{Partial}: Learn half the melody (restores at half effectiveness). \textit{Fail}: The keystone cracks further --- Roller Integrity $-1$. \\
  \hline
  \textbf{Seek the Rain Queen's blessing} & Resolve + Survival (DV 7) & \textit{Success}: Gain a \textbf{Rain Charm} (once per scene, ignore one weather hazard). \textit{Partial}: Gain a charm that breaks after one use. \textit{Fail}: Offend the queen --- Ancestor Favor $-2$. \\
  \hline
  \end{longtable}
  
  \vspace{2mm}
  \subsection*{The Pursuing Threat}
  
  On the third day, the party discovers what has been following the city: a \textbf{Dust Walker} --- a spirit of abandoned keystones, drawn to cities that have lost their way. It is the size of a man but made of swirling red dust, with hollow eyes that weep sand. It does not speak. It does not need to. It simply follows, counting the days.
  
  \vspace{2mm}
  \textbf{Dust Walker --- Phase Timer [6]}
  \begin{itemize}
  \item At 2 segments: The dust grows thick. All navigation rolls at $-1$ die.
  \item At 4 segments: The Walker appears at the edge of camp. It points at a party member. That member gains the \textbf{Marked} condition (the Walker will target them first in a confrontation).
  \item At 6 segments: The Walker attacks. Treat as a boss with \textbf{Phases}: Counting $\rightarrow$ Marking $\rightarrow$ Claiming. \textbf{Cracks}: Offer it a forgotten keystone, speak the name of a city that chose to stop, or let it witness a true act of ancestral honor.
  \end{itemize}
  
  \vspace{2mm}
  \subsection*{Root-binder Negotiation}
  
  The Root-binders are led by a woman named \textbf{Nomvula's Shadow} (no relation to the Rain Queen --- she took the name in defiance). She offers a deal:
  
  \textit{``Let the city stop. Plant acorns in the rollers. The forest will take the stone. Your people will become tree-kin. You will never hunger again. The Hollow does not walk among roots.''}
  
  The party can:
  \begin{itemize}
  \item \textbf{Accept} --- The city stops. The party gains the \textbf{Green Pact} (permanent +1 die to survival in forests, but the Hollow's attention increases by 1 each session). The Mason-Kin will never forgive you.
  \item \textbf{Refuse} --- Root-binder Attention ticks +2 immediately. The Root-binders will sabotage one roller per day until the timer fills.
  \item \textbf{Bargain} --- Offer a Keystone Charter or a Stone-Song Fragment. The Root-binders will delay their attack for 3 days and offer one \textbf{Green Blessing} (a single use of their acorn magic).
  \end{itemize}
  
  \vspace{2mm}
  \subsection*{Ancestor Appeasement}
  
  The ancestors are angry because the city's last mason, a woman named \textbf{Phendu}, died without passing on the stone-song. Her ghost haunts the keystone chamber, singing the melody but forgetting the final note.
  
  To appease the ancestors, the party must:
  \begin{enumerate}
  \item \textbf{Find Phendu's apprentice} --- A one-eyed woman named \textbf{Sola}, now living as a Root-binder outcast. She knows the final note but refuses to sing it because the Mason-Kin exiled her.
  \item \textbf{Convince Sola to return} --- Requires a \textbf{Presence + Sway test (DV 6)}. On success, she returns and teaches the final note. On failure, she demands a price: the party must cut a promise to the Mason-Kin (lose one Keystone Charter or equivalent).
  \item \textbf{Sing the full stone-song at the keystone} --- Requires a \textbf{Spirit + Performance test (DV 5)}. On success, Ancestor Favor fills completely, and the keystone cracks heal. On failure, the ancestors are offended --- Roller Integrity $-2$ immediately.
  \end{enumerate}
  
  \vspace{2mm}
  \subsection*{Possible Outcomes}
  
  \begin{longtable}{|p{0.25\textwidth}|p{0.55\textwidth}|}
  \hline
  \textbf{Outcome} & \textbf{Consequences} \\
  \hline
  \textbf{City saved (new roller)} & Khomani's Progress survives. The party gains a \textbf{Keystone Charter} (right to dwell in any Ngomebe city for one season) and the title \textit{Stone-Friend}. \\
  \hline
  \textbf{City saved (stone-song healed)} & The original roller is restored. The party gains a \textbf{Stone-Song Fragment} (permanent +1 die to negotiate with Ngomebe). The ancestors whisper your names. \\
  \hline
  \textbf{City stops, but survivors evacuate} & The city becomes a Still City --- a hollow shell. The Hollow comes. The party escapes with a \textbf{Dead Keystone Shard} (can be traded to the Root-binders or used to bargain with the Dust Walker). The Mason-Kin blame you. \\
  \hline
  \textbf{City stops, party joins Root-binders} & The party becomes tree-kin. Gain \textbf{Green Pact} (as above). The Mason-Kin become enemies. The Hollow notices you more often. \\
  \hline
  \textbf{Root-binders destroyed} & The groves burn. The party gains the enmity of all Root-binders (they will sabotage any city you travel with). Gain a \textbf{Mason-Kin Honor Blade} (+1 die to combat in stone cities). \\
  \hline
  \textbf{Dust Walker banished} & The spirit retreats. Gain a \textbf{Dust Walker's Eye} (a gem that reveals the location of any abandoned keystone within a day's travel). \\
  \hline
  \end{longtable}
  
  \vspace{2mm}
  \subsection*{Reward Table}
  
  \begin{longtable}{|p{0.3\textwidth}|p{0.55\textwidth}|}
  \hline
  \textbf{Reward} & \textbf{Effect} \\
  \hline
  Keystone Charter & Right to dwell in any Ngomebe city for one season. Can be renewed by service. \\
  \hline
  Stone-Song Fragment & Reroll one failed Survival or Craft test per session. \\
  \hline
  Green Pact (Root-binder) & +1 die to survival in forests; Hollow's Attention +1 per session. \\
  \hline
  Mason-Kin Honor Blade & +1 die to combat when fighting in or near a walking city. \\
  \hline
  Dust Walker's Eye & Once per journey, locate the nearest abandoned keystone. \\
  \hline
  Rain Charm & Once per scene, ignore one weather hazard. Breaks after 3 uses. \\
  \hline
  \end{longtable}
  
\end{tcolorbox}

% Index entries
\index{Ngomebe|see{Stone Men}}
\index{Theme generator}
\index{Stone Men generator}
\index{Places!Ngomebe}
\index{People!Ngomebe}
\index{Complications!Ngomebe}
\index{Rewards!Ngomebe}
\index{Mason-Kin}
\index{Root-binder}
\index{Aeler engineers}
\index{Keystone}
\index{Walking city}
\index{First City}
\index{Still Voice}
\index{Ox-Master}
\index{Stone-song}
\index{Sledge-wax}
\index{Roller-bearing}
\index{Ancestral Healers}
\index{Izinyanga}
\index{Tokoloshe}
\index{Impundulu}
\index{Rain Queen}
\index{Umbhemisi}
\index{Local Patrons!Ngomebe}