\section{Nihori --- ``The Isles of the Dawn Spirit''}
\label{chap:nihori}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{Duel at dawn. Honor the storm. The ninth wave never breaks.}

\begin{quote}
  \textbf{Storm Emperor's Ghost (Elite):}  
  ``I ruled the waves before the Hollow learned to walk. My empire sank because I spoke a name I should not have. Do not speak my name. Do not ask what it was. The sea remembers. That is enough.''
\end{quote}

\begin{quote}
  \textbf{Wanderer (Commoner):}  
  ``I carry a broken sword and a grudge. The sword was my father's. The grudge is mine. The man who killed him died nine years ago. I still walk the isles, looking for his ghost. The ghost is looking for me too.''
\end{quote}

\textbf{Genre / Mood:} Storm-tossed honor duels, kami politics, oni cults, volcanic doom, drowned ghosts, yokai in the ash.

\textbf{Starting Location:} A bamboo grove on a clifftop, where two duelists face each other at dawn. Ash from the volcano dusts their shoulders. A Shrine Maiden rings a copper bell: ``The Shogun's envoy demands the duel stop. The daimyo of the southern isle says it will not. The storm-spirit is watching. Choose a side --- or the ninth wave will choose for you.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Nihori is a chain of volcanic islands, shrouded in mist and lashed by storms --- a shattered shogunate where the old empire sank beneath the waves and the \textbf{Lethai-al} guard the drowned kingdom's memory. The people are wanderers, duelists, shrine maidens, and oni-touched outcasts. They worship the storm spirits (kami of wind, wave, and crater) and fear the ninth wave. The Hollow breathes in the salt foam; Oath-Keepers ride the tide. The Shogun sits in a half-ruined castle, his authority challenged by three rival daimyo who sharpen their blades and their treaties. \textbf{Oni cults} whisper in the ash wastes, offering forgotten bargains. \textbf{Y{\=u}rei} (vengeful ghosts) wail from sunken ships. \textbf{Kappa} drag children from tidal pools. \textbf{Tengu} mock travelers from the high crags. The Storm Emperor's ghost does not rule --- but his silence haunts every island. The ninth taboo is absolute: never speak the name of the drowned capital. The sea will hear you, and it will ask why you called.

\vspace{2mm}
\textbf{What you notice first:} The smell of salt and volcanic ash. The way every traveler carries a sword, even the tea-sellers. The constant sound of bells on shrine gates --- ringing in the wind. And the silence when a daimyo's retinue passes. The distant wail of a \textit{funay{\=u}rei} (a drowned sailor's ghost) calling for a boat that never comes.

\textbf{The one rule that kills newcomers:} Never draw steel in a shrine. The kami will witness. They may not forgive.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Wave and the Edge of the Blade}

Power here is measured in duels won, storms survived, and the number of clans that bow to your banner. The Shogun rules from Castle Kurotani, but his writ fades as the islands stretch into the mist. Three daimyo vie for supremacy: the \textbf{Northern Crane} (old blood, slow to anger, quicker to negotiate), the \textbf{Southern Serpent} (pragmatic, pirates in his pay, a smile like a cracked bell), and the \textbf{Western Oni} (rumored to have made a pact with a Crater Oni, his eyes glow faintly red in the dark). Between them, the \textbf{Shrine Maidens} keep the peace --- not through steel, but through storm-calling and the threat of the ninth wave. A duel is not merely a fight; it is a negotiation sanctified by kami. First blood decides the truth. The loser yields a story. The winner returns a cup of sake --- and a prayer at the nearest shrine.

\subsection*{Prominent NPCs of Nihori}
\label{sec:nihori-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Shogun Atsumori & Figurehead ruler of the shattered realm & Keep the daimyo from open war & He has not spoken in seven years. His attendants speak for him. \\
\hline
Daimyo Sadako (Southern Serpent) & Ruler of the southern isles, pirate ally & Break the Shogun's peace and claim the throne & She wears a necklace of brass hooks --- one for each ship she has taken. \\
\hline
Daimyo Yoritomo (Northern Crane) & Old-blood traditionalist & Restore the shogunate's authority by the old laws & He has never lost a duel. He has never drawn first blood. \\
\hline
The Oni-Eyed (Daimyo of the West) & Ruler of the western ash wastes, rumored oni-touched & Ascend to a new form --- neither human nor spirit & His eyes glow red in darkness. His retainers wear masks. \\
\hline
Shrine Maiden Hana & Keeper of the Storm Shrine & Pacify the spirit of the ninth wave & She rings a bell that can only be heard underwater. \\
\hline
Wanderer Ronin & Masterless swordsman & Find the ghost of his father's killer & He has never lost a duel. He has also never drawn first blood. \\
\hline
Pirate Captain Red-Hook & Kahfagian raider & Control the strait between the isles & His left hand is a brass hook. The hook is also a key. \\
\hline
The Drowned Scholar (Ghost) & A Lethai-al who died before completing his research & Ask the living to retrieve a scroll from the sunken capital & He appears at low tide, holding a skeletal oar. \\
\hline
Oni-Brewer (Cult Leader) & A priest of the Crater Oni who brews memory-erasing sake & Turn the daimyo's court into puppets & His breath smells of ash and plum wine. He never blinks. \\
\hline
Kappa-Binder & A hermit who catches kappa in clay pots & Trade kappa secrets for human memories & He has a bowl-shaped dent in his head --- a kappa's mark. \\
\hline
\end{longtable}

\subsubsection*{Daimyo Strings (Quick Reference)}
\begin{itemize}
  \item \textbf{Southern Serpent (Sadako):} Can provide pirate escorts, forged papers, or a safe harbor --- for a price in favors or a future raid. \textit{Typical String:} ``A ship that will ignore customs once.''
  \item \textbf{Northern Crane (Yoritomo):} Can offer legal arbitration, access to the archive temple, or a duel champion --- expects honor and a public oath in return. \textit{Typical String:} ``A witnessed verdict that clears your name.''
  \item \textbf{Oni-Eyed:} Can grant oni-touched powers (a die bonus to intimidation, darkvision, etc.) --- but accepting gives $+1$ \textit{Oni Corruption} immediately. \textit{Typical String:} ``A promise made to the Crater Oni --- it will be settled in memory or blood.''
\end{itemize}

\subsection*{Names of Nihori}
\label{sec:nihori-names}

Inspired by the northern isles: short, with vowel endings and occasional glottal stops. Surnames denote a clan, a wave, or a drowned ancestor.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/Epithet} & \textbf{Wave-Name} \\
\hline
Kenji & Hana & Wave-walker & \textit{Ninth Breath} \\
\hline
Ronan & Yuki & Storm-kin & \textit{Salt-Tongue} \\
\hline
Kael & Mira & Drowned-line & \textit{Unforgotten} \\
\hline
Takeshi & Sora & Foam-born & \textit{Tide-Charm} \\
\hline
Jiro & Nami & Ancestor-salt & \textit{the Unshadowed} \\
\hline
Yoshio & Rin & Kappa-binder & \textit{Bowl-Head} \\
\hline
\end{longtable}

\paragraph*{(Forge/Grove/Shrine)} A swordsmith's forge with a waterfall cooling tank; a bamboo grove where duels are fought at dawn; a storm-shrine on a cliff with ropes for climbing.

\section*{Spades --- Places}
\label{sec:nihori-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Swordsmith's Forge (Volcano-Fired)} --- A waterfall cooling tank; prayer strips on the bellows.
  \hfill \textit{The swordsmith prays before each strike. The prayer is a verse from the old epic.} \texttt{[CRAFT]}
\item \textbf{Bamboo Grove (Duel Ground)} --- A clearing where the stalks are scarred by blade cuts.
  \hfill \textit{The bamboo bleeds sap when a duel is dishonorable. The sap is red.} \texttt{[HONOR]}
\item \textbf{Storm-Shrine (Cliff)} --- A wooden platform on a sea cliff, with ropes for climbing.
  \hfill \textit{The ropes are braided with human hair. The hair is from sailors who survived shipwrecks.} \texttt{[KAMI]}
\item \textbf{Drowned Palace (Ruins)} --- Half-submerged columns visible at low tide.
  \hfill \textit{The columns are carved with names. The names are the drowned kingdom's royalty.} \texttt{[DEATH]}
\item \textbf{Volcano Shrine (Dormant)} --- A crater where the forbidden name was spoken.
  \hfill \textit{The crater is warm. Speaking any name aloud here causes the ground to tremble.} \texttt{[ONI]}
\item \textbf{Castle Kurotani (Shogun's Seat)} --- A half-ruined fortress on a clifftop, roofs patched with salvaged ship timber.
  \hfill \textit{The Shogun's throne is cracked. The crack is shaped like a wave.} \texttt{[POLITICS]}
\item \textbf{Southern Serpent's Fleet Anchorage} --- A hidden harbor where pirate dhows and war junks moor together.
  \hfill \textit{The flags are all different. The loyalty is all to the same purse.} \texttt{[AUTHORITY]}
\item \textbf{Oni's Caldera (Western Ash Wastes)} --- A crater lake where the Oni-Eyed daimyo holds court.
  \hfill \textit{The water is warm and red. It tastes of iron and forgotten promises.} \texttt{[CORRUPTION]}
\item \textbf{Crane Clan's Archive Temple} --- A library built into a sea cliff, scrolls sealed in wax.
  \hfill \textit{The clan has not opened the oldest wing in a century. The seal is still warm.} \texttt{[TRADITION]}
\item[J] \textbf{The Sunken Gate} --- An underwater archway that leads to the drowned capital.
  \hfill \textit{It can only be reached by holding your breath for nine heartbeats.} \texttt{[THRESHOLD]}
\item[Q] \textbf{Storm-Reader's Lookout} --- A wooden tower on the highest cliff.
  \hfill \textit{From here, the Storm-Reader can see three storms coming. He has never been wrong.} \texttt{[WEATHER]}
\item[K] \textbf{The Ninth Wave's Rock} --- A black stone that the ninth wave strikes every full moon.
  \hfill \textit{The wave leaves behind a single white shell. The shell contains a memory.} \texttt{[OMEN]}
\item[A] \textbf{The Drowned Capital (Myth)} --- A sunken city said to rise once every century.
  \hfill \textit{Those who enter can speak to the dead. The price is a breath.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Wanderer/Maiden/Pirate)} Wanderer with a broken sword; Shrine Maiden with a copper bell; Pirate Captain with a brass hook.

\section*{Hearts --- People \& Factions}
\label{sec:nihori-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Wanderer (Ronin)} --- A man with no master, a scarred blade, and a cup of sake.
  \hfill \textit{He offers a cup to every opponent. ``Drink,'' he says. ``The duel will be fair.''} \texttt{[HONOR]}
\item \textbf{Shrine Maiden Hana} --- A woman with a copper bell on a rope. She rings it at dawn.
  \hfill \textit{The bell's tone tells the fishing fleet whether to sail. It has never been wrong.} \texttt{[KAMI]}
\item \textbf{Pirate Captain Red-Hook} --- A woman with a brass hook for a hand. She lost the original to a sea monster.
  \hfill \textit{The hook is also a key. It opens a chest of treasure maps --- and a dossier on the Southern Serpent's spies.} \texttt{[STEALTH]}
\item \textbf{Storm-Reader} --- Gray-haired, eyes the color of storm clouds. He can feel the pressure change in his bones.
  \hfill \textit{``The storm will arrive in nine hours,'' he says. ``Bring your offerings --- or your prayers.''} \texttt{[WEATHER]}
\item \textbf{Swordsmith} --- A woman with calloused hands and a prayer strip on her forehead.
  \hfill \textit{She names each sword after a dead warrior. The sword remembers. The warrior's ghost may visit.} \texttt{[CRAFT]}
\item \textbf{Shogun's Chamberlain} --- A soft-spoken man with a fan painted with a crane. He speaks for the silent Shogun.
  \hfill \textit{He never raises his voice. The daimyo lower theirs when he enters.} \texttt{[POLITICS]}
\item \textbf{Daimyo Sadako (Southern Serpent)} --- A woman with a scarred lip, a fleet of dhows, and a network of informants.
  \hfill \textit{She wears a necklace of brass hooks --- one for each ship she has taken. She adds a new one each season.} \texttt{[AUTHORITY]}
\item \textbf{Daimyo Yoritomo (Northern Crane)} --- An old man with perfect posture, a blade that has never drawn blood, and a library of old laws.
  \hfill \textit{He has never lost a duel. He has never drawn first blood. His opponents always apologize.} \texttt{[TRADITION]}
\item \textbf{The Oni-Eyed} --- A figure in a tattered mask, one eye glowing red. He rules the ash wastes from a throne of cooled lava.
  \hfill \textit{His retainers wear masks of their own faces --- cast from the moment they swore loyalty.} \texttt{[CORRUPTION]}
\item[J] \textbf{Oni-Brewer (Cult Leader)} --- A bald man with burn scars, a ceramic sake jug, and a sweet, rotten smile.
  \hfill \textit{His sake erases memories. He sells it to daimyo who want to forget old burdens --- and to assassins who want to forget guilt.} \texttt{[BURDEN]}
  \textit{Mechanic: Drinking the Oni-Brewer's sake clears one Obligation segment or eases one Burden --- but you also lose one random minor memory (a face, a password, a skill point for one scene). The GM records what you forgot; it may return as a nightmare or a ghost.}
\item[Q] \textbf{Storm Emperor's Ghost} --- A column of steam over the water, shaped like a crowned man.
  \hfill \textit{He asks: ``Has anyone spoken the name?'' If you say yes, he weeps. If you say no, he waits.}
  \textit{Triggered event: When the ninth wave strikes or the Sunken Gate opens, the ghost solidifies into a towering figure of foam and lightning. He points at one character. That character hears the name of the drowned capital whispered in their ear --- and now the Hollow knows they know.} \texttt{[DEATH]}
\item[K] \textbf{Crater Oni (Sealed)} --- A horned shadow visible only in the reflection of volcanic vents.
  \hfill \textit{It offers bargains: a year of luck for a memory, a century of life for a death. The Oni-Eyed accepted. He does not remember the price.} \texttt{[PATRON]}
\item[A] \textbf{The First Drowned (Legend)} --- The first Lethai-al who died in the sinking. She sits at the bottom of the sea.
  \hfill \textit{If you bring her a memory, she will teach you to breathe water. If you bring her a lie, she will drown you.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Ashfall/Duel/Pirate)} Ashfall from the volcano --- all actions Desperate; a duel challenge is issued --- refusal loses face; a pirate flotilla claims the strait --- pay toll or fight.

\section*{Clubs --- Complications/Threats}
\label{sec:nihori-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Ashfall (Volcano)} --- The dormant crater emits ash. Visibility drops. All actions are Desperate. \texttt{[WEATHER]}
  \hfill \textit{The ash is warm. It smells of old names. Speaking a name makes the ash thicken --- and the oni listen.}
\item \textbf{Duel Challenge (Issued)} --- A Lethai-al challenges a party member to a duel. Refusal loses face. \texttt{[HONOR]} \texttt{[COMBAT]}
  \hfill \textit{The duel is to first blood. The loser yields a story. The story must be true. The winner returns a prayer.}
\item \textbf{Pirate Flotilla (Southern Serpent's Fleet)} --- Kahfagian pirates block the strait. They demand a toll of one tide-charm. \texttt{[AUTHORITY]} \texttt{[COMBAT]}
  \hfill \textit{The charm must be from a drowned ancestor. If the party has no tide-charm, they must offer a memory or a promise instead --- the pirates are not picky.}
\item \textbf{Oni Cult Initiation} --- A traveler is marked by the Oni-Eyed's cult. Their shadow has horns. \texttt{[CORRUPTION]} \texttt{[FEAR]}
  \hfill \textit{The cult demands a service: steal a scroll from the Crane Clan's archive, or the mark will spread.}
\item \textbf{Tide-Charm Stolen} --- Someone steals a pouch of charms. The ancestors are angry. \texttt{[DEATH]} \texttt{[OBLIGATION]}
  \hfill \textit{The thief is a foreign merchant in the pay of the Southern Serpent. The charms are in a cargo hold guarded by oni-touched pirates.}
\item \textbf{Daimyo Summit (War Threatened)} --- The three daimyo agree to meet. The Shogun's peace holds --- barely. \texttt{[POLITICS]}
  \hfill \textit{Any false move could ignite a civil war. The party is there as witnesses --- or scapegoats.}
\item \textbf{Volcano Name Whispered} --- Someone speaks the forbidden name near the crater. The ground shakes. \texttt{[DEATH]} \texttt{[DISASTER]}
  \hfill \textit{The speaker is a child of the Oni-Eyed's household. The child did not know. The volcano remembers. The oni laughs.}
\item \textbf{Ninth Wave Warning} --- The full moon is tomorrow. The ninth wave will be larger than any in a century. \texttt{[WEATHER]}
  \hfill \textit{The fishing fleet cannot be recalled. Someone must warn them --- or bargain with the storm-spirit for their lives.}
\item \textbf{Sake House Brawl (Oni-Brewer's Sake)} --- A dispute over a duel's fairness turns violent. The ninth cup is poured. \texttt{[SOCIAL]} \texttt{[OBLIGATION]}
  \hfill \textit{The ninth cup is for the dead. Someone has just died. The brawl is now a blood feud --- and everyone has forgotten why.}
\item[J] \textbf{Sunken Gate Opens} --- The underwater archway becomes passable. Something swims out. \texttt{[THRESHOLD]} \texttt{[DEATH]}
  \hfill \textit{It is a drowned spirit. It wants its name back. It will trade a secret for it --- or take yours.}
\item[Q] \textbf{Storm-Reader's Silence} --- The old man stops reading the storm. The fishermen do not know where to sail. \texttt{[WEATHER]} \texttt{[GRIEF]}
  \hfill \textit{He is grieving. His son died at sea. He will not speak until the body is found --- or until someone tells him a true story of the sea.}
\item[K] \textbf{Duel to the Death (Oni-Eyed's Assassin)} --- A duel that was meant to be to first blood becomes a killing. \texttt{[HONOR]} \texttt{[TRAP]}
  \hfill \textit{The killer is a spy from the Western Ash Wastes. The duel was an assassination. The party is blamed --- unless they can prove the oni-touched intervened.}
\item[A] \textbf{The Ninth Wave Takes} --- The massive wave crashes over the cliff shrine. The bells are silent. \texttt{[DEATH]} \texttt{[CATASTROPHE]}
  \hfill \textit{The shrine maiden is gone. The storm-spirit is free. The Hollow walks the shore. The only way to appease it is to speak the name of the drowned capital --- and accept the consequences.}
\end{enumerate}

\paragraph*{(Token/Bell/Charm)} Swordsmith's token (one reroll on a failed melee attack); shrine bell (reveals hidden spirits for one scene); wanderer's oath-bracelet (one use to compel a service).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:nihori-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Swordsmith's Token (Iron)} --- A small iron disc. Reroll a failed melee attack once. \texttt{[COMBAT]}
  \hfill \textit{The token is stamped with a prayer. The prayer is for a clean cut. The swordsmith made it for a warrior who never returned.}
\item \textbf{Shrine Bell (Copper)} --- A small copper bell. Ring it to reveal hidden spirits for one scene. \texttt{[KAMI]}
  \hfill \textit{The bell's tone is inaudible to humans. The spirits hear it clearly. They may answer --- or they may hide.}
\item \textbf{Wanderer's Oath-Bracelet (Leather)} --- A braided leather bracelet. Demand a service from any wanderer once. \texttt{[HONOR]}
  \hfill \textit{The bracelet is stained with sake. The wanderer cannot refuse. The service is one night's vigil or one day's guide.}
\item \textbf{Storm-Charm (Bone)} --- A carved bone. Ignore one weather penalty on a journey leg. \texttt{[WEATHER]}
  \hfill \textit{The bone is from a sailor who died in a storm. His spirit protects you --- but he will ask for a story in return.}
\item \textbf{Tide-Charm (Shell)} --- A white shell. Whisper a drowned ancestor's name to calm a storm for one hour. \texttt{[DEATH]} \texttt{[WEATHER]}
  \hfill \textit{The shell is warm. It glows when the ancestor is listening. Do not lie. They know.}
\item \textbf{Duelist's Sake Cup} --- A ceramic cup used in a fair duel. Demand a duel to first blood. \texttt{[HONOR]}
  \hfill \textit{The cup is chipped. The chip is from a blade that drew first blood. The loser's story is carved into the rim.}
\item \textbf{Pirate Captain's Hook (Brass)} --- A brass hook prosthetic. Use it as a key to open any chest in the cove. \texttt{[STEALTH]}
  \hfill \textit{The hook is also a weapon. It can cut rope and bone. It also fits the lock on the Southern Serpent's war chest.}
\item \textbf{Volcano Priest's Whisper} --- A single word: the forbidden name. Speak it to cause a tremor. \texttt{[DEATH]} \texttt{[MAGIC]}
  \hfill \textit{The word burns the tongue. Use once. The volcano will remember. The oni will hear.}
\item \textbf{Drowned Scholar's Scroll (Waterproof)} --- A scroll retrieved from the sunken capital. Contains a forgotten secret. \texttt{[LEGEND]}
  \hfill \textit{The scroll is wet. It will never dry. The secret is about the Storm Emperor's true name --- and why he spoke it.}
\item[J] \textbf{Ninth Wave's Shell} --- A white shell from the Ninth Wave's Rock. Break it to experience a memory. \texttt{[OMEN]}
  \hfill \textit{The memory is always a death. The death is always a drowning. The memory may be yours --- or an ancestor's.}
\item[Q] \textbf{Storm-Reader's Flute (Bone)} --- A bone flute. Play it to predict the weather for the next 24 hours. \texttt{[WEATHER]}
  \hfill \textit{The tune is sad. It always predicts a storm. It is always right. The Storm-Reader will not explain why.}
\item[K] \textbf{Sunken Gate Key (Coral)} --- A piece of coral shaped like a key. Opens the Sunken Gate. \texttt{[THRESHOLD]}
  \hfill \textit{The key is alive. It grows one size each time it is used. After nine uses, it becomes a coral door that cannot be moved --- and the gate is permanently open.}
\item[A] \textbf{First Drowned's Breath (Glass Vial)} --- Seawater from the bottom of the sunken capital. Drink it to breathe water for one hour. \texttt{[LEGEND]}
  \hfill \textit{The water tastes of salt and sorrow. The lungs will ache. The memory of drowning will stay with you.}
\end{enumerate}

% ============================================================
% ADDITIONAL GENERATORS
% ============================================================

\subsection*{Storm-Spirit Negotiation (d6 Outcomes)}
\label{sec:nihori-storm-spirit}

When the party attempts to bargain with a storm kami (at a shrine, during a typhoon, or at the Ninth Wave's Rock), roll on this table to determine the outcome. The GM sets the initial DV based on the offering (a tide-charm, a story, or a cup of sake).

\begin{longtable}{|p{2cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{Outcome} & \textbf{Result} \\
\hline
\textbf{Critical Success} & The spirit grants a \textit{Storm-Charm} (ignore one weather penalty on any future voyage). The kami may also whisper a warning about the next oni cult activity. \\
\hline
\textbf{Success} & The spirit calms the storm for the party only. Other ships in the area are not protected. Gain $+1$ \textit{Kami Favor}. \\
\hline
\textbf{Partial} & The storm abates, but the spirit demands a tide-charm as a toll. Gain a Burden Timer [4]. If you cannot pay, the storm returns in 1d4 hours. \\
\hline
\textbf{Failure} & The spirit is offended. All sailing rolls are Desperate for the next leg. The kami may send a \textit{kappa} or \textit{funay{\=u}rei} (drowned ghost) to harass the party. \\
\hline
\textbf{Critical Failure} & The ninth wave rises. Test Body + Athletics (DV 5) or be swept overboard. Lose 1 Supply and gain the \textit{Marked by the Storm} condition (weather hazards are one step worse for the next journey). \\
\hline
\end{longtable}

\subsection*{Oni Cult Whispers (d8)}
\label{sec:nihori-oni-whispers}

When the party encounters the influence of the Oni-Eyed or the Crater Oni, roll on this table for cryptic messages, omens, or cult demands.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d8} & \textbf{Whisper} \\
\hline
1 & ``The Oni-Eyed seeks a name. Give him yours, and he will give you a ship --- and a shadow that walks behind you.'' \\
\hline
2 & ``The Crater Oni accepts memories in trade. Your first kiss. Your mother's face. A fair trade --- until the price deepens.'' \\
\hline
3 & ``The ninth wave is the Oni's breath. When it breaks, the Hollow listens. Those who hear it must speak a true name or lose one of their own.'' \\
\hline
4 & ``The Sunken Gate opens at low tide. Bring a drowned scholar's scroll --- or a life. The gate does not care which.'' \\
\hline
5 & ``The Shogun is a ghost. The daimyo know. They are afraid to admit it. The Oni-Eyed will laugh when the truth speaks itself.'' \\
\hline
6 & ``The Southern Serpent's fleet carries oni-touched pirates. Their shadows have horns. Their captain has no reflection.'' \\
\hline
7 & ``The Storm Emperor's ghost speaks only to those who have lost something at sea. What did you lose? The oni can help you forget it.'' \\
\hline
8 & ``The oni-brewer's sake lifts burdens. It also lifts the one who carries them. Choose carefully. The Crater Oni is patient --- and thirsty.'' \\
\hline
\end{longtable}

\subsection*{Duel Complications (d6)}
\label{sec:nihori-duel-complications}

When a formal duel is underway, roll on this table for unexpected twists that can change the outcome or escalate the stakes.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d6} & \textbf{Complication} \\
\hline
1 & \textbf{Kami Witness} --- A shrine kami manifests as a whirlwind or a flash of lightning. Both duelists must test Resolve (DV 4) or lose their turn. The kami favors the one who offers a prayer. \\
\hline
2 & \textbf{Interrupted by Ashfall} --- Volcanic ash darkens the sky. Ranged attacks (including thrown weapons) suffer $-1$ die. The duel may be postponed unless both parties agree to continue. \\
\hline
3 & \textbf{Oni Interference} --- A horned shadow flickers at the edge of the grove. The duelist with higher \textit{Oni Corruption} gains $+1$ die to intimidation but $-1$ die to honor-based rolls. The oni is betting on the outcome. \\
\hline
4 & \textbf{The Ninth Echo} --- The clashing of blades rings nine times. On the ninth ring, the ground trembles. Both duelists must test Body + Athletics (DV 3) or stumble, losing the next exchange. \\
\hline
5 & \textbf{A Witness's Secret} --- A bystander reveals that one duelist is under an oni bargain. The duel can be halted for a \textit{Geas Test} (Spirit + Lore vs. DV 4). Success cancels the bargain for the duration of the duel; failure binds the bargainer further. \\
\hline
6 & \textbf{The Loser's Story} --- Instead of first blood, the loser must tell a true story of a broken promise. If the story moves the kami, the duel ends in a draw and both parties gain $+1$ \textit{Kami Favor}. If the story is false, the loser takes Harm 1 (spiritual backlash). \\
\hline
\end{longtable}

\subsection*{Kami Favor Boons (d6)}
\label{sec:nihori-kami-boons}

When a character spends \textit{Kami Favor} (see Additional Features), roll or choose a boon from this table. Each boon costs 1--3 Favor as indicated.

\begin{longtable}{|p{2cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{Boon (Cost)} & \textbf{Effect} \\
\hline
\textbf{Calm Seas (1 Favor)} & The kami calms the waters for one leg. All sailing rolls gain $+1$ Position. \\
\hline
\textbf{Wind at Back (1 Favor)} & Your ship gains $+1$ Speed for one leg. On a critical success, you arrive one segment early. \\
\hline
\textbf{Kami's Warning (2 Favor)} & Before the next storm, you receive a vision. The GM must answer one yes/no question about the weather or a sea hazard. \\
\hline
\textbf{Duel Blessing (2 Favor)} & In your next duel, you gain $+1$ die to Honor-based actions (e.g., showing mercy, refusing a dirty trick). \\
\hline
\textbf{Ancestor's Tide (3 Favor)} & A drowned ancestor guides your ship through a reef or shoal. Ignore one \textit{Sunken Gate} or \textit{Pirate Flotilla} complication. \\
\hline
\textbf{Oni Ward (3 Favor)} & For the remainder of the current scene, you are immune to oni-touched fear effects. Your shadow does not grow horns even if your Corruption track is high. \\
\hline
\end{longtable}

\subsection*{Yokai Encounters (d8)}
\label{sec:nihori-yokai}

When traveling through the wilder parts of Nihori (ash wastes, deep forests, forgotten shores), roll on this table for a supernatural encounter. Yokai are not always hostile --- they are tricksters, testers, and reminders of a world older than the shogunate.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d8} & \textbf{Encounter} \\
\hline
1 & \textbf{Kappa in the Ford} --- A river kappa blocks a shallow crossing. It demands a cucumber, a bow, or a memory. Refuse, and it will try to pull a party member underwater (Grapple, Body+Endurance DV 3 to resist). If you win, it may grant a \textit{Kappa's Secret} --- $+1$ die on all Swimming or Sailing rolls for the rest of the journey. \\
\hline
2 & \textbf{Floating Lantern (Funay{\=u}rei)} --- A drowned sailor's ghost drifts toward your boat, holding a paper lantern. It asks, ``Where is my grave?'' If you answer truthfully (a specific cove or shrine), it blesses your voyage ($+1$ Position on the next sea leg). If you lie, it capsizes your boat. \\
\hline
3 & \textbf{Tengu on the Pass} --- A long-nosed tengu sits on a stone, blocking the mountain trail. It challenges you to a game of riddles or a contest of pride. If you win, it gives you a \textit{Tengu Fan} (once, create a gust of wind to extinguish fires or fill sails). If you lose, it steals your name --- you forget one Bond or Skill for 1d4 scenes. \\
\hline
4 & \textbf{Nure-onna (Wet Woman)} --- A serpent-woman with the head of a beautiful woman and the body of a snake appears on the shore. She offers to heal a wound or ease one Burden --- in exchange for a vial of your blood or a lock of your hair. Refusing angers her (Harm 1, poison). Accepting gives you a \textit{Promise Token} she will call in at the worst moment. \\
\hline
5 & \textbf{Jor{\=o}gumo (Spider Bride)} --- A woman in silk robes invites you to her teahouse. The teahouse is a web. She offers you a deal: information about the Oni-Eyed's location in exchange for a memory of a living family member. If you refuse, she attacks (Cap 3, venomous). If you accept, you gain the information --- but your family member will dream of spiders for a week. \\
\hline
6 & \textbf{Yuki-onna (Snow Woman)} --- During a sudden cold snap, a woman in white appears. She asks for a fire or a warm cloak. If you give it freely, she grants you a \textit{Frost Charm} (once, ignore a heat-based hazard). If you refuse, she freezes your breath --- you gain 1 Fatigue and suffer $-1$ die on all physical rolls for the next scene. \\
\hline
7 & \textbf{Onmoraki (Birth-Death Bird)} --- A bird with a human face hatches from a corpse. It speaks in riddles about a death that will occur in the next 24 hours. If the party prevents that death, the Onmoraki becomes a guardian spirit (Cap 1 follower for one session). If they fail, it curses them --- the next time anyone rolls a 1, it becomes a 10 (but the GM gains 2 SB). \\
\hline
8 & \textbf{The Sh{\=o}j{\=o} (Drunkard Sage)} --- A red-faced, singing creature floats on a sake barrel offshore. It offers to share its brew. Drinking it gives $+2$ dice to all social rolls for one scene --- but you cannot lie, and every truth you speak is heard by the nearest oni cult spy. The Sh{\=o}j{\=o} laughs and vanishes, leaving an empty barrel. \\
\hline
\end{longtable}

\subsection*{Haunted Blades (Cursed Swords)}
\label{sec:nihori-haunted-blades}

In Nihori, every sword has a spirit, and some swords have ghosts. A character who acquires a masterwork blade (cost 8 XP) may choose to make it \textit{haunted}. The blade grants a powerful benefit but also a curse.

\begin{longtable}{|p{2cm}|p{5cm}|p{5cm}|}
\hline
\rowcolor{gray!20}
\textbf{Blade} & \textbf{Benefit} & \textbf{Curse} \\
\hline
\textbf{Whisper of the Drowned} & $+1$ die to attack rolls against pirates and oni. & The blade whispers the names of drowning sailors at night. You cannot recover Fatigue from sleep while at sea. \\
\hline
\textbf{Storm's Edge} & Once per scene, as a free action, gain $+1$ Effect on a melee attack during a storm. & The blade attracts lightning. On a Critical Failure, you are struck for Harm 1 (shock). \\
\hline
\textbf{Ancestor's Tooth} & The blade can cut through spirit wards. Ignore 1 point of Armor from ghosts and oni. & The ancestor's ghost appears in your dreams, demanding you avenge a centuries-old wrong. Start an \textit{Ancestor's Claim} timer [6]. \\
\hline
\textbf{Broken Vow (Cursed)} & $+2$ dice on the first attack of any duel. & If you lose a duel, the blade breaks --- and you cannot repair it without a quest to the sunken capital. \\
\hline
\end{longtable}

\section*{Quick use notes}
\label{sec:nihori-quick-use}
\begin{itemize}
\item Draw until you have all four suits. Highest rank sets main Timer size: 2--5 $\rightarrow$ 4 segments, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10.
\item Diamonds are codified outcomes (tokens, bells, charms) that change position rather than call for a roll.
\item If any A appears, echo \textbf{sea-omens}---a wave that breaks twice, a bell that rings underwater, a name that echoes from the foam.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:nihori-features}

\begin{itemize}
\item \textbf{Duel Law:} Disputes in Nihori are settled by duel to first blood. The loser yields a story; the winner returns a cup of sake. A duel can be witnessed by anyone; the witness's word is law. Refusing a duel is an admission of guilt. In the shogunate courts, a duel verdict can overturn a written agreement.
\item \textbf{Storm Season Timer [8]:} Tracks the approach of the monsoon. When it fills, a typhoon strikes --- all travel becomes Desperate for 2 legs. The timer ticks each day during storm season. The Shrine Maidens can delay it by one day for a suitable offering (a tide-charm, a memory, or a promise).
\item \textbf{Kami Favor (Reputation Track):} Instead of debt, track \textbf{Kami Favor [6]}. Gain favor by respecting shrines, winning duels honorably, and appeasing storm-spirits. Spend favor to ask a kami for a boon (see Kami Favor Boons table). At 0 favor, shrines are hostile --- all social rolls at shrines are Desperate.
\item \textbf{Oni Corruption Track [6]:} Represents the influence of the Oni-Eyed's cult and the Crater Oni. Tick when a character makes a bargain with an oni, drinks oni-brewed sake, or speaks the forbidden name. At 3 segments: the character's shadow grows horns. At 6: the character becomes oni-touched --- they gain $+1$ die to intimidation but suffer $-1$ die to all social rolls with shrine maidens and can never enter a shrine without being attacked.
\item \textbf{Shogunate Faction Timers:} Track the three daimyo's schemes:
  \begin{itemize}
    \item \textbf{Southern Serpent's Ambition [6]} --- advances when she bribes officials, builds her fleet, or sponsors pirate raids. When full, she declares open war on the Shogun.
    \item \textbf{Northern Crane's Patience [6]} --- retreats when he is humiliated or his allies defect. When it reaches 0, he abandons the Shogun's peace and marches south.
    \item \textbf{Oni-Eyed's Ascension [8]} --- advances each time a character accepts an oni bargain or the party fails to stop a cult ritual. When full, the Crater Oni manifests in the material world.
  \end{itemize}
\item \textbf{Yokai and the Old Exchange:} Unlike simple monsters, yokai in Nihori treat with travelers. They may offer bargains, demand tolls, or curse those who break taboos. Use the Yokai Encounters table to introduce them, and track any promises made to a yokai as a \textit{Yokai Promise} timer [4]. When it fills, the yokai returns to call it in.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Storm Emperor's Ghost} --- Khemesh + The Witness. \textit{The pressure of a forbidden name, the memory of a drowned kingdom.}
    \item \textbf{The Crater Oni} --- Malachai. \textit{Cursed gifts, bargains that leave scars, power that transforms you from the inside.}
    \item \textbf{The Ninth Wave} --- Ra{\'e}yn. \textit{The storm that does not break, the tide that remembers every sailor.}
    \item \textbf{The Bell of Drowned Shrines} --- The Sealed Gate. \textit{The boundary between the living and the dead, guarded by bells and salt.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Drowned Capital and the Kalrantha Amulet (Dhahara):} The sunken city is said to contain a temple that holds a fragment of the demon Varash. The Veiled Sisters want it. The Oni-Eyed wants to bargain with it. The Shrine Maidens want it to stay drowned.
  \item \textbf{Tide-Charms and the Tulkani (Way of Silk):} Tulkani story-tellers sometimes carry Lethai-al tide-charms as proof of a witnessed drowning. A charm from Nihori is worth twice as much --- and attracts twice as much Hollow attention.
  \item \textbf{Storm-Spirits and the Veiled Sisters (Ashaan):} The Veiled Sisters have tried to bind storm-spirits for use in their fleet. The Lethai-al have foiled nine attempts. The tenth attempt is being funded by the Southern Serpent's pirate gold.
  \item \textbf{Oni Bargains and the Chain (Bitter Root):} Daughters of the Cord refuse to treat with oni --- the price is always paid in souls, and the Cord cannot collect from a damned spirit. An oni-touched Daughter is a contradiction; one may be hiding in the ash wastes, trying to untie her cord from a Crater Oni's bargain.
  \item \textbf{The Ninth Wave and the Hollow (Eastern Realms):} Aelaerem hedge-keepers say that the ninth wave is the Hollow's breath. When it crashes, the Hollow exhales. Those who hear it must speak a true name or lose one of their own.
\end{itemize}

\subsection*{Nihori Superstitions (burn for $+1$ die once)}
\begin{itemize}
  \item \textbf{Bread to the Foam:} Crumble a crust at the water's edge. Ignore the first \emph{Pirate Flotilla} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Rope:} Tie a single knot in a shrine climbing rope. Cancel \emph{Storm-Spirit Offended} or take $-1$ die on your next sailing roll (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Wave:} Whisper a drowned ancestor's name into the ninth wave. Gain $+1$ Effect when navigating a storm this scene, but mark \emph{Obligation} to the Hollow. \texttt{[SUPERSTITION]}
  \item \textbf{Sake to the Oni:} Pour a cup of sake onto volcanic stone. Cancel one tick of the \emph{Oni Corruption} track --- but the oni will remember your face. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:nihori-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{clan retainers or shrine maidens}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{duelists or pirate lookouts}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{oni cultists or the Storm-Reader's spies}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}.
\item \textbf{Supply Strain:} Mark $+1$ segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local daimyo's guard or shrine keeper} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{ashfall or rising water}.
\item \textbf{Foe Appears:} An \textbf{oni-touched ronin or a pirate boarding party} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{oni cultists, pirate ships, or clan samurai} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes---\textbf{duel becomes a killing / volcano trembles / ninth wave crests early}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Shogun's Chamberlain, Oni-Eyed, or Storm Emperor's Ghost} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically---\textbf{volcano erupts / tsunami strikes / shrine collapses}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice---\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Nihori (Duel Law):} A witness recants, the first blood is disputed, a duelist's story is revealed as false.
\item \textbf{Nihori (Storm \& Kami):} A kami demands a toll, a storm shifts direction, the ninth wave's timing changes.
\item \textbf{Nihori (Oni Cult):} A shadow grows horns, a memory curdles, an oni bargain's fine print appears.
\item \textbf{Nihori (Yokai):} A kappa laughs from a riverbank, a tengu's feather falls on your shoulder, a snow woman's breath fogs your vision.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Navigation):} On a hit, shift one band. On a strong hit, also impose \emph{Fog Bank}.
\item \textbf{Cut the Duel (Presence + Honor):} On a hit, freeze range for one exchange; if a duelist's sake cup is in your possession, $+1$ die.
\item \textbf{Weather the Wave (Resolve + Survival):} On a hit in \emph{Ninth Wave Warning}, you pick who keeps their footing; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Honor \& Oni)}
\begin{itemize}
\item \textbf{The Oni-Eyed (Ascended Cultist)} --- \emph{Phases}: Whispered Bargain $\rightarrow$ Horned Shadow $\rightarrow$ Oni Manifest. \emph{Cracks}: speak his original name, destroy the mask, offer a memory the oni cannot refuse.
\item \textbf{The Storm Emperor's Ghost (Drowned Sovereign)} --- \emph{Phases}: Silent Tide $\rightarrow$ Wave of Names $\rightarrow$ The Ninth Breaks. \emph{Cracks}: speak the forbidden name at the Sunken Gate, return the drowned scholar's scroll, let the Storm-Reader negotiate.
\item \textbf{Southern Serpent (Pirate Admiral)} --- \emph{Phases}: Fleet Gathers $\rightarrow$ Strait Blockade $\rightarrow$ Throne Claim. \emph{Cracks}: expose her spy network, sink her flagship, convince the Crane to ally with the Shogun.
\item \textbf{The Kappa King (Lord of the Freshwater Depths)} --- \emph{Phases}: Drown the Ford $\rightarrow$ Steal the Children $\rightarrow$ Claim the Well. \emph{Cracks}: return a stolen name, offer a cucumber of great value, survive his grip and pull off his bowl (draining his power).
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In Nihori, power flows through duels, storms, the favor of kami, and the dangerous bargains of oni. The Shogun is a figurehead; the three daimyo are the real players. A clan chief's authority rests on how many duels he has won and how many shrines he has funded. A shrine maiden's on how many storms she has predicted and how many oni she has banished. Yokai watch from the margins, ready to lend aid or call in a promise.

\textbf{For the GM:}  
Use the \textbf{Storm Season Clock} as a visible threat. Let \textbf{Duel Law} be a way to resolve conflicts without combat --- or a way to escalate them when honor is insulted. Make \textbf{Tide-Charms} valuable but dangerous --- each use may attract the Hollow or the attention of an oni. Track the three daimyo's \textbf{Faction Timers} as a campaign front; the party's actions can accelerate or delay civil war. Use \textbf{Kami Favor} as a measure of standing --- players must choose which shrines to support and which kami to appease. The \textbf{Storm-Spirit Negotiation} and \textbf{Oni Cult Whispers} tables add depth to supernatural encounters, while \textbf{Duel Complications} keep combat fresh and honor-bound. Introduce \textbf{Yokai Encounters} to make the wilderness feel alive and capricious, and offer \textbf{Haunted Blades} to characters willing to accept a curse for power.
\end{tcolorbox}

% Index entries
\index{Nihori|see{Storm Isles}}
\index{Theme generator}
\index{Nihori generator}
\index{Places!Nihori}
\index{People!Nihori}
\index{Complications!Nihori}
\index{Rewards!Nihori}
\index{Duel law}
\index{Kami favor}
\index{Oni corruption}
\index{Storm season}
\index{Tide-charms}
\index{Shogunate factions}
\index{Crane clan}
\index{Serpent clan}
\index{Oni-Eyed}
\index{Storm Emperor's ghost}
\index{Crater Oni}
\index{Ninth wave}
\index{Yokai}
\index{Kappa}
\index{Tengu}
\index{Funay{\=u}rei}
\index{Jor{\=o}gumo}
\index{Yuki-onna}
\index{Onmoraki}
\index{Haunted blades}
\index{Local Patrons!Nihori}