\section{Dhahara: The Elephant Kingdoms and the Iron Passes}
\label{chap:dhahara}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The elephant remembers every path. The mountain counts every breath. The peacock watches the ninth throne.}

\begin{quote}
  \textbf{Maharaja of Ahimsa (Elite):}  
  ``The Peacock Throne is not a seat. It is a promise made by the first sun king: that the nine kingdoms would remain nine, never one, never fewer. I do not rule. I balance. A peacock's tail has nine eyes. Each eye watches a different kingdom. When one closes, the throne weeps. The Aeler in the high passes are not our masters. They are our debtors. The mountain remembers who first carved the tunnels.''
\end{quote}

\begin{quote}
  \textbf{Viceroy Thrain (Aeler Elite):}  
  ``The passes do not care for your lineage or your gods. They care for iron, tea, and the weight of a sworn oath. Aeler and Dhaharan have bled into the same stone here. I wear a dwarf's breath-counter and a Dhaharan fire-token. The mountain accepts both. The lowland kings may dream of unity, but the iron does not dream. It only waits.''
\end{quote}

\begin{quote}
  \textbf{Elephant Mahout (Commoner):}  
  ``My grandfather drove elephants for the Sun Kings. My father drove them for the Aeler. I drive them for the Viceroy. The elephants do not care who holds the howdah. They care that the pass is clear and the fodder is fresh. They also remember the old songs — the ones about the ninth kingdom that never was.''
\end{quote}

\textbf{Genre / Mood:} Elephant warfare, monsoon ritual, honour as ancestral dharma, high‑altitude survival horror, industrial colonialism.

\textbf{Starting Location:} The Peacock Throne Hall in Ahimsa, where a golden peacock with ruby eyes watches the court. The Maharaja's chamberlain approaches: ``The Maharaja wishes to renew the Nine-Kingdom Pact. But the King of Vajrapur has sent a forged genealogy. The Queen of Mahishmati has refused to attend. And the ninth kingdom — the one that is always missing — has sent a messenger. The messenger is a ghost. Meanwhile, Viceroy Thrain has sent a breath-counter and a demand: the iron tithe is due. Speak carefully. The peacock is watching.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Between the high passes of the Aeler viceroyalty and the Brass Coast lies a fertile, contested lowland: the land of the Nine Kingdoms, the heart of Dhahara. Here, ancient dynasties still claim descent from the sun, and the Peacock Throne of Ahimsa casts its shadow over half a dozen rival courts. Rivers are borders; elephants are war machines; every harvest is a negotiation between kings who have never met but whose grandfathers swore blood oaths. The Covenant has a strong presence in the sacred cities; the Aeler have counting houses in the iron towns; the Sidhi sail up the rivers to trade. But Dhahara belongs to the Dhaharan — to their poets, their temple dancers, their bandit kings, and their queens who rule from behind silk screens.

North, in the high passes, the Aeler viceroys rule on behalf of the distant High King. Here, the air is thin and cold, the rivers run white with snowmelt, and the passes are measured in breath‑counts rather than miles. Iron mines scar the hillsides, worked by Dhaharan labourers and overseen by Aeler engineers. Elephant caravans haul ore down to the Brass Coast; mule trains carry silver and tea up to the mountain holds. The Aeler have established a viceroyalty here — not conquest, but a long lease, renewed every nine years. The local rajas pay tribute in iron and receive protection from lowland enemies. It is an uneasy peace, written in bell‑codes and toll‑stones. The architecture blends Aeler stonework with Dhaharan domes; the cuisine mixes dwarf‑bread with spiced tea; the law is a creole of Covenant statutes and Aeler breath‑bonds.

But both north and south share a common truth: the number ``nine'' is a wound — and a promise. There have never been nine kingdoms. The number is a ritual, a prayer that the Hollow will not notice that one of the nine is always missing. That missing kingdom changes. Sometimes it is a real kingdom that fell to plague or war; sometimes it is a kingdom that never was, a placeholder kept empty to balance the cosmic scales. The Peacock Throne weeps because the missing kingdom is always the one that mattered — and because the Hollow collects the weight of its absence. In the high passes, the ninth breath is the one the mountain takes. In the lowlands, the ninth throne is the one the Hollow watches. North and south, the obligation is the same.

\vspace{2mm}
\textbf{What you notice first:} The smell of jasmine and dust, and the ringing of iron hammers. The way every crossroads has a painted stone — a boundary marker, a warning, a prayer. The constant murmur of bazaars where haggling is a form of theatre. The way every noble touches their left wrist when speaking of ancestors — a gesture of honour. The silence after a bell rings in the mountains — the silence is for the mountain to answer.

\textbf{The one rule that kills newcomers:} Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one — a breath, a memory, a shadow. In the lowlands, never praise a king's elephant. The elephant is sacred; your praise will be taken as an insult to the king's own mount. Compliment the mahout instead. The mahout will tell the elephant.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Elephant and the Gaze of the Hollow}

Power here is measured in river rights, elephant herds, the thickness of a dynasty's archive, and the number of breaths a miner can hold. In the lowlands, a king's authority rests on how many generations back to the sun — but also on how many of those generations kept their promises. The Peacock Throne of Ahimsa is the nominal centre, but the other eight kings vie for influence, and the balance shifts with every monsoon. The Covenant judges provide legal arbitration; the temple dancers control the ritual calendar; the bandit kings of Kalinga pay no tribute to anyone. In the high passes, the Viceroy answers to the Aeler High King, but also to the local rajas, the Covenant judges, and the elephant clans. Every decision is a negotiation between mountain custom and lowland law. The Aeler engineers speak a creole of their own tongue; the Dhaharan nobles wear Aeler breath-counters as jewellery; and the elephants — the elephants remember who fed them during the famine of the ninth lease.

And somewhere, in the gap between the eighth and the ninth — between the eighth breath and the ninth breath, between the eighth throne and the ninth throne — the Hollow waits to collect what the missing kingdom owes.

\subsection*{Prominent NPCs of Dhahara}
\label{sec:dhahara-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Maharaja Suryavarman & King of Ahimsa (Peacock Throne) & Unite the nine under his dynasty & His pet peacock follows him; its cry is a warning of assassination attempts. \\
\hline
Viceroy Thrain & Aeler governor of the Iron Passes & Maximise ore output for the High King & He wears a bell around his neck; it rings when he lies. \\
\hline
Rani Chandravati & Queen of Vajrapur (Iron Kingdom) & Protect the iron mines from Aeler encroachment & She wears a bracelet of dwarf teeth — each tooth is from an engineer she outwitted. \\
\hline
King Bhim & Buffalo King of Mahishmati & Expand his pastoral kingdom into Anga & He has never lost a buffalo race; his buffalo is named ``Conqueror.'' \\
\hline
Covenant High Judge Vasu & Head of the Kosala sacred circuit & Enforce contract law across the nine kingdoms & He carries a scale that never balances; he says justice is never equal, only weighed. \\
\hline
Temple Dancer Mira & Devadasi of the Peacock Throne & Keep the peace through ritual & Her dance is said to bring rain; she has stopped three droughts. She dreams of the missing ninth. \\
\hline
Mine-Captain Korr & Aeler engineer in the high passes & Keep the tunnels safe and the quotas met & He counts his breaths in eights; he has never reached nine. \\
\hline
Rani Jaya & Local raja of the eastern hills (high passes) & Balance tribute to Aeler with loyalty to her clan & She lost an arm to a mine collapse; the prosthetic is a pickaxe. \\
\hline
Prince Karna of Anga & Heir to the eastern fortress & Recover the lost tunnels of Anga & He walks with a limp from a childhood tiger attack; the tiger was his mother's pet. \\
\hline
Bandit Queen Jaya & Ruler of the Kalinga hills & Carve a tenth kingdom from the margins & Her parrot is a transformed spy; it still works for three masters. \\
\hline
Scribe Govinda & Master of the Silent Court & Keep the Covenant's secrets from leaking & He writes with a quill made from a dead man's finger — the finger of a forger. \\
\hline
Tarkan & Elephant caravan master (high passes) & Move iron to the coast before the snows close the passes & He speaks to his elephants in Dhaharan; they answer in Aeler. \\
\hline
The Ninth King's Herald (Ghost) & A translucent messenger who appears at every succession crisis & Remind the living that the missing kingdom still has a vote & It speaks in the voice of the last king of the lost ninth. That voice changes each time. \\
\hline
\end{longtable}

\subsection*{Names of Dhahara}
\label{sec:dhahara-names}

Sanskrit and Prakrit roots, with honourifics (Rao, Singh, Deva, Bai). In the high passes, Aeler clan names and Dhaharan hill titles mix. Surnames often denote clan, profession, or a specific mine. A character's full name includes an honour‑syllable (\textit{-rinn}, \textit{-datt}, \textit{-hrin}), spoken only when swearing oaths.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/Surname} & \textbf{Epithets} \\
\hline
Surya & Chandravati & Somavamsi & \textit{the Peacock} \\
\hline
Bhim & Mira & Kalachuri & \textit{Iron-Fisted} \\
\hline
Vasu & Ganga & Yadava & \textit{Buffalo-Loyal} \\
\hline
Rajendra & Lakshmi & Gupta & \textit{the Coin-Eater} \\
\hline
Krishna & Durga & Pala & \textit{the Learned} \\
\hline
Karna & Jaya & Malla & \textit{the Unseen} \\
\hline
Govinda & Tara & Shailendra & \textit{the Silent} \\
\hline
Thrain & Jaya & Ironfoot & \textit{the Bell-Breaker} \\
\hline
Korr & Gaura & Stonehand & \textit{Breath-Counter} \\
\hline
Borin & Meera & Vikram & \textit{the Elephant's Friend} \\
\hline
\end{longtable}

\paragraph*{(Lowland)} Hall of the Peacock Throne, Iron Gate Market, Buffalo Race Track, Sacred City of Sravasti, Anga Fort, Western Port of Surashtra, Rajagriha Ruins, Elephant Stables of Ahimsa, Well of the Nine Kings, Forgotten Throne of Kalinga, Silent Court of Kosala, Peacock Gardens, Whispering Pass, Nine Kings' Tomb.

\paragraph*{(Highland)} Iron Mine of the Crying Face, Aeler Toll Bridge, Viceroy's Hall, Elephant Corral (Mountain), Silent Pass, Mine of the Ninth Seal, Tea Terraces (Aeler-Dhaharan), Dwarf Counting-House, Raja's Fort (Hill), Bridge of Nine Echoes, Abandoned Smelter, Viceroy's Archive, First Lease Stone.

\section*{Spades — Lowland Places (Nine Kingdoms)}
\label{sec:dhahara-lowland-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peacock Throne Hall} --- A vast chamber with pillars of carved marble; the throne is a peacock of gold and lapis lazuli. `[AUTHORITY]`
  \hfill \textit{The peacock's eyes are rubies. They follow anyone who approaches the throne. The throne has never been sat upon by a usurper. The ninth eye is missing — it fell out during the last succession war and was never replaced.}
\item \textbf{Iron Gate Market (Vajrapur)} --- A bazaar outside the fortress; the smell of iron dust and spices. `[TRADE]`
  \hfill \textit{The market has a section where slaves are sold. The slaves wear iron collars stamped with the king's mark. One collar is always empty — reserved for the ghost of the missing ninth kingdom.}
\item \textbf{Buffalo Race Track (Mahishmati)} --- A dusty oval; the betting is fierce. `[SPORT]`
  \hfill \textit{The track has a shrine at the centre — a buffalo that never lost. Its skeleton is on display. The skeleton sometimes snorts. When it does, a promise is about to come due.}
\item \textbf{Sacred City of Sravasti (Kosala)} --- A walled city with a thousand temples; the Covenant courthouse is here. `[JUSTICE]`
  \hfill \textit{The city has no bells. The Covenant forbids them. The silence is enforced by armed judges. The silence is loudest at noon. The ninth temple has no door.}
\item \textbf{Anga Fort} --- A hilltop fortress with a dry moat; the walls are painted with tiger stripes. `[DEFENSE]`
  \hfill \textit{The fortress has never been taken. The king's secret is a network of hidden tunnels. The tunnels have not been mapped in a century. One tunnel leads to the missing ninth kingdom — or so the ghosts say.}
\item \textbf{Western Port of Surashtra} --- A Sidhi trading post; the harbour is full of dhows. `[TRADE]`
  \hfill \textit{The port has a lighthouse shaped like a woman. The woman is a Sidhi saint. Her lamp is lit with green flame. The flame flickers when a ninth kingdom obligation is recalled.}
\item \textbf{Rajagriha Ruins} --- The former imperial capital, now a graveyard of broken pillars. `[RUINS]` `[DEATH]`
  \hfill \textit{The ruins are haunted by the ghosts of Gupta emperors. They ask for news of their empire. They do not accept bad news. They also ask who sits on the ninth throne.}
\item \textbf{Elephant Stables of Ahimsa} --- A complex of stone stalls; each elephant has its own name carved above the door. `[COMMUNITY]`
  \hfill \textit{The names are not the elephants' birth names. They are the names of enemies they trampled. The elephants remember. One stall is always empty — for the elephant that carried the ninth king.}
\item \textbf{Well of the Nine Kings} --- A deep stepwell with nine landings; each landing has a statue of a different king. `[MYSTERY]`
  \hfill \textit{The water is black. It shows the face of the next king to die. The face changes when you blink. The ninth landing shows only a reflection of the viewer — the missing king is always you.}
\item \textbf{Forgotten Throne of Kalinga} --- A stone chair sunk in a marsh; the Bandit Queen holds court there. `[SECRET]`
  \hfill \textit{The throne was once a king's. He lost a bet. The bet was his kingdom. The marsh still smells of that laughter. The throne has a ninth armrest, worn smooth by no hand.}
\item[J] \textbf{Silent Court of Kosala} --- A Covenant courthouse with no speaking allowed; all arguments are written. `[JUSTICE]`
  \hfill \textit{The judges write their verdicts in disappearing ink. Only they know the secret reagent. The reagent is tears. The ninth judge has no face — only a veil of grey silk.}
\item[Q] \textbf{Peacock Gardens} --- A walled garden behind the throne hall; only the king may enter. `[LEGEND]`
  \hfill \textit{The garden contains a single peacock that is immortal. It was the first king's familiar. It has not eaten in three hundred years. It has nine feathers in its tail. One falls each time a kingdom dies.}
\item[K] \textbf{Whispering Pass (Anga-Mahishmati)} --- A narrow defile where the wind carries conversations. `[SPY]`
  \hfill \textit{Spies use it to listen to enemy camps. The pass is named for the dead — they whisper too. Their whispers are always true. The ninth whisper is always a name you have never heard.}
\item[A] \textbf{Nine Kings' Tomb} --- A subterranean complex where the founders of each dynasty are buried. `[DEATH]` `[LEGEND]`
  \hfill \textit{The tomb is sealed. The key is a single word that changes each year — the name of the current weakest king. The word is never spoken aloud. The ninth sarcophagus is empty. It waits.}
\end{enumerate}

\section*{Spades — Highland Places (Mountain Satrapies)}
\label{sec:dhahara-highland-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Iron Mine of the Crying Face} --- A deep shaft with a carved face above the entrance.
  \hfill \textit{The face weeps rust. The tears are iron oxide. The mountain is mourning its dead.} `[BREATH]`
\item \textbf{Aeler Toll Bridge} --- A stone arch over a gorge; a dwarf with a scale sits at the far end.
  \hfill \textit{The toll is not in coin — it is in weight. Your truth is weighed against a bar of iron.} `[TOLL]`
\item \textbf{Viceroy's Hall} --- A longhouse of timber and stone; a throne of iron ore.
  \hfill \textit{The throne is cold. The viceroy never sits on it — only stands behind it.} `[LEASE]`
\item \textbf{Elephant Corral (Mountain)} --- A fenced area with heavy chains; elephants are chained by one leg.
  \hfill \textit{The chains have names — each is the elephant's first kill.} `[COMMUNITY]`
\item \textbf{Silent Pass} --- A high mountain pass where no bird flies.
  \hfill \textit{The air is too thin for sound. Arguments must be written in the snow.} `[BREATH]`
\item \textbf{Mine of the Ninth Seal} --- A sealed tunnel with a Covenant seal on the timbers.
  \hfill \textit{Inside is a demon that the miners woke. The seal is cracking.} `[MAGIC]` `[LEASE]`
\item \textbf{Tea Terraces (Aeler-Dhaharan)} --- Stone-step fields where tea shrubs grow.
  \hfill \textit{Each terrace is a different clan's share of the harvest. The tea tastes of slate.} `[TRADE]`
\item \textbf{Dwarf Counting-House} --- A low building with brass scales and an iron record.
  \hfill \textit{The record tracks breath‑bonds. It weighs the air.} `[TOLL]`
\item \textbf{Raja's Fort (Hill)} --- A stone fortress with painted gates and a single tower.
  \hfill \textit{The tower is also a fire shrine. The flame is lit with dwarf-coal.} `[COMMUNITY]`
\item[J] \textbf{Bridge of Nine Echoes} --- A rope bridge over a chasm; each step echoes nine times.
  \hfill \textit{The ninth echo is always a whisper. It is the name of the next person to fall.} `[OMEN]`
\item[Q] \textbf{Abandoned Smelter} --- A ruined furnace where iron was once refined.
  \hfill \textit{The slag heaps are still warm. Something lives in the warmth.} `[MAGIC]`
\item[K] \textbf{Viceroy's Archive} --- A subterranean vault of treaties and toll receipts.
  \hfill \textit{The archive is sealed with Aeler keystone wax and Dhaharan clay.} `[LEASE]`
\item[A] \textbf{The First Lease Stone} --- A monolith carved with the original treaty between Aeler and Dhahara.
  \hfill \textit{The stone is warm. It glows when the lease is renewed. It cracks when the lease is broken.} `[LEASE]` `[OMEN]`
\end{enumerate}

\paragraph*{(Lowland NPCs)} Peacock Throne chamberlain, iron mines overseer (dwarf), buffalo herder, Covenant judge (High Judge Vasu), temple dancer (Mira), bandit of Kalinga, spice merchant (Sidhi), poet of the Peacock Throne, well keeper of the Nine Kings, war elephant mahout, ghost of Rajagriha, silent judge's clerk, peacock keeper, ninth king's ghost.

\paragraph*{(Highland NPCs)} Mine-captain (Aeler), Rani Jaya, Covenant judge (Sister Gaura), elephant caravan master (Tarkan), Aeler toll-keeper, smelter worker, treasure hunter, raja's spymaster, tea trader (Fhara), blind pilgrim, ghost of the landslide, viceroy's cipher clerk, demon's whisper (from the Ninth Seal).

\section*{Hearts — Lowland People \& Factions}
\label{sec:dhahara-lowland-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peacock Throne Chamberlain} --- A eunuch in silk; he controls access to the king. `[AUTHORITY]`
  \hfill \textit{He has a mirror on a stick; he uses it to see around corners. He has seen nine assassins before they struck. He has not warned the king about the tenth. The tenth works for the missing kingdom.}
\item \textbf{Iron Mines Overseer (Dwarf)} --- Wears an Aeler breath-counter. He is loyal to the queen, but his cousin is a viceroy's clerk. `[POLITICS]`
  \hfill \textit{He keeps two records. One is for the queen. One is for the Aeler. Both are forgeries. The truth is in his head. He is losing his memory. The ninth entry in every record is always wrong.}
\item \textbf{Buffalo Herder} --- A young man with a staff and a red turban. `[COMMUNITY]`
  \hfill \textit{He can find water in a drought by listening to the ground. His buffalo have never gone thirsty. He shares his wells with travellers. The ninth well he dug produced blood instead of water. He does not speak of it.}
\item \textbf{Covenant Judge (High Judge Vasu)} --- A grey-haired man with a brass scale. He has never smiled. `[JUSTICE]`
  \hfill \textit{His clerks have never seen him eat. He drinks only rainwater. He says food clouds the mind. His judgments are always delivered at noon. The ninth judgment he ever made was never recorded. It was for the missing kingdom.}
\item \textbf{Temple Dancer (Mira)} --- A woman with anklets of silver and eyes lined with kohl. `[RITUAL]`
  \hfill \textit{Her dance is a prayer. The prayer is for peace. She has never failed — but the peace never lasts. She dances anyway. In the ninth step of her dance, she always stumbles slightly. No one knows why.}
\item \textbf{Bandit of Kalinga} --- A man with a parrot on his shoulder, a scar across his face. `[STEALTH]`
  \hfill \textit{The parrot is his conscience. It screams when he lies. He lies often. The parrot has lost its voice. It regained it once, to name the ninth king's murderer. The name was never spoken aloud.}
\item \textbf{Spice Merchant (Sidhi)} --- A Sidhi trader with a leather bag full of cinnamon. He speaks seven languages. `[TRADE]`
  \hfill \textit{He lies in all of them. His weights are true, though. He says the weights are the only honest thing about him. The ninth weight in his set is a forgery — it is exactly one grain lighter. He uses it to pay obligations to the Hollow.}
\item \textbf{Poet of the Peacock Throne} --- A woman with a stringed instrument, her voice cracked from singing. `[SOCIAL]`
  \hfill \textit{She writes verses for the king. The verses are encoded warnings. The king cannot read. He has her sing them instead. Her ninth verse is a curse that activates when the missing kingdom returns. She prays it never does.}
\item \textbf{Well Keeper of the Nine Kings} --- A blind woman who sits by the stepwell. `[MYSTERY]`
  \hfill \textit{She knows the depth of the water by the echo of a dropped stone. She knows the future by the same echo. She does not charge for her visions. She charges for the silence afterwards. The ninth echo always says the same word: “Wait.”}
\item \textbf{War Elephant Mahout} --- An old man who has never lost a battle. He whispers to his elephant in a language of clicks. `[COMBAT]`
  \hfill \textit{The elephant's name is “Sorry.” It has trampled nine enemies. It always apologises. The ninth apology was accepted by a ghost. The ghost now rides beside it.}
\item[J] \textbf{Ghost of Rajagriha} --- A translucent figure in a crown. He asks one question: “Does my empire still stand?” `[DEATH]`
  \hfill \textit{The answer must be a lie, or he will weep and flood the ruins. The lie must be kind. He can tell the difference. He has asked the question nine times. The ninth answer was silence. He is still waiting.}
\item[Q] \textbf{Silent Judge's Clerk} --- A young woman who writes in disappearing ink. She is the only one who knows the reagent. `[JUSTICE]`
  \hfill \textit{She has been attacked twice. She carries a dagger made of frozen ink. The ink melts after one use. The ninth time she uses it, the dagger will not melt. She does not know what that means.}
\item[K] \textbf{Peacock Keeper} --- A boy who tends the immortal peacock. The peacock speaks to him. `[LEGEND]`
  \hfill \textit{It tells him which courtiers are plotting. The boy has never been wrong. He does not tell the king. He tells the peacock. The peacock has nine warnings stored. It will speak the ninth when the missing kingdom returns.}
\item[A] \textbf{Ninth King's Ghost} --- A shade that walks the tomb. It points at the killer — but the killer is always a living king. `[DEATH]` `[FEAR]`
  \hfill \textit{It points at whoever is closest. That person is accused. The accusation is always false. The crowd does not care. The ninth time it points, it will point at the true killer. No one will believe it.}
\end{enumerate}

\section*{Hearts — Highland People \& Factions}
\label{sec:dhahara-highland-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Mine-Captain (Aeler)} --- A dwarf with a tuning fork and a level. He can hear a crack in a beam from a hundred paces.
  \hfill \textit{He never walks under a keystone. The mountain has ears.} `[BREATH]`
\item \textbf{Rani Jaya} --- A young woman in armour, riding a caparisoned elephant. She lost an arm to a mine collapse.
  \hfill \textit{The prosthetic is a pickaxe. She can write with it. The letters are iron filings.} `[COMMUNITY]`
\item \textbf{Covenant Judge (Sister Gaura)} --- Robed in grey, a water-clock on her belt. The clock runs on tears.
  \hfill \textit{She refills it from her own eyes when she hands down a death sentence.} `[LEASE]`
\item \textbf{Elephant Caravan Master (Tarkan)} --- Old, missing two fingers, tattooed with elephant footprints.
  \hfill \textit{He knows the name of every elephant that has crossed the pass in fifty years.} `[COMMUNITY]`
\item \textbf{Aeler Toll-Keeper} --- A dwarf with a brass scale and a record. He weighs your goods and your lies.
  \hfill \textit{The scale tips for both. A lie makes the pan dip lower.} `[TOLL]`
\item \textbf{Smelter Worker} --- A young man with red-rimmed eyes. He coughs iron dust.
  \hfill \textit{His spittle stains the ground. He will not live past thirty. He knows.} `[LABOR]`
\item \textbf{Treasure Hunter} --- A foreigner digging in an old landslide. He is looking for the tomb of a prince.
  \hfill \textit{The prince is the one who stole the Kalrantha amulet. The treasure hunter is a fool.} `[MAGIC]`
\item \textbf{Raja's Spymaster} --- A woman with a limp; she uses a walking stick that is also a blowpipe.
  \hfill \textit{The darts are tipped with tea — caffeine that stops hearts.} `[POLITICS]`
\item \textbf{Tea Trader (Fhara)} --- A merchant with a caravan of mules. He sells tea to the Aeler; they pay in iron.
  \hfill \textit{He sells iron to the coast; they pay in silver. He is the richest man in the passes.} `[TRADE]`
\item[J] \textbf{Blind Pilgrim} --- A woman in white, walking the passes alone. She is searching for a shrine where a bell rang without a clapper.
  \hfill \textit{The shrine is a cave. The cave is the mine of the ninth seal.} `[FAITH]`
\item[Q] \textbf{Ghost of the Landslide} --- A figure in grey, seen only at dusk. It points at the pass.
  \hfill \textit{It is pointing at a safe route — or a trap. No one has survived to find out.} `[OMEN]`
\item[K] \textbf{Viceroy's Cipher Clerk} --- A dwarf who writes diplomatic letters in code.
  \hfill \textit{The code is a poem. The poem is about a lost mine. The mine is real.} `[LEASE]`
\item[A] \textbf{Demon's Whisper (from the Ninth Seal)} --- A voice heard in the sealed mine.
  \hfill \textit{It offers a deal: a year of good fortune for a single memory. The memory is the last breath of a dying friend.} `[MAGIC]` `[LEASE]`
\end{enumerate}

\paragraph*{(Lowland Complications)} Elephant stampede — trained to kill; buffalo race riot — crowd turns violent; Covenant excommunication — contract voided, goods seized; temple dance cursed — the rain does not come; spy in the Peacock Gardens; Nine Kings' Tomb robbers; Well of Nine Kings poisoned; Peacock Throne assassination attempt; Silent Court leak; royal genealogy forged; Whispering Pass ambush; immortal peacock's cry; ghost's demand; well's vision.

\paragraph*{(Highland Complications)} Cave-In (sabotage); toll dispute; ninth echo on the Bridge of Nine Echoes; landslide; breath shortage; spy in camp; Covenant audit; Aeler bell code change; elephant sickness (curse); Viceroy's summons; Silent Pass's cold; demon wakes (Ninth Seal); lease stone cracks.

\section*{Clubs — Lowland Complications/Threats}
\label{sec:dhahara-lowland-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Elephant Stampede (Royal)} --- The king's elephants break loose; they are trained to kill. `[DISASTER]` `[COMBAT]`
  \hfill \textit{The stampede was ordered by a rival queen. The target is a visiting diplomat. The diplomat is you. The elephants are not confused. One elephant wears a saddle from the missing ninth kingdom.}
\item \textbf{Buffalo Race Riot} --- A disputed finish; the crowd turns violent. `[SOCIAL]` `[COMBAT]`
  \hfill \textit{The dispute is a trap. The losing king wants an excuse to invade. You are the excuse. The buffalo are innocent. The ninth buffalo in the race has no shadow.}
\item \textbf{Covenant Excommunication} --- High Judge Vasu declares your contract void; your goods are seized. `[JUSTICE]` `[WEIGHT]`
  \hfill \textit{The excommunication is a bribe. The bribe was paid by a merchant you undercut. The merchant is in the crowd, smiling. He wears a ninth ring on a hand that has only eight fingers.}
\item \textbf{Temple Dance Cursed} --- Mira stumbles; the rain does not come. The priest blames the party. `[RITUAL]` `[SOCIAL]`
  \hfill \textit{The sacrifice is a goat. The goat is a spy. The spy is working for Vajrapur. The goat will not stop talking. It speaks of the ninth kingdom in a language that sounds like bells.}
\item \textbf{Spy in the Peacock Gardens} --- Someone overheard the king's war plans. `[POLITICS]` `[STEALTH]`
  \hfill \textit{The spy is a poet. The poet is you — you were hired to write verses, not to eavesdrop. The king has a long memory for verse. The verse you overheard was the ninth stanza of a forbidden epic.}
\item \textbf{Nine Kings' Tomb Robbers} --- Grave robbers break into the tomb. They seek the key word. `[DEATH]` `[THEFT]`
  \hfill \textit{The word is “forgiveness.” The robbers are agents of a rival kingdom. They do not know the word. They are digging anyway. The ninth sarcophagus is open. It is empty. The robbers did not open it.}
\item \textbf{Well of Nine Kings Poisoned} --- The black water turns red; people sicken. `[DEATH]` `[WEATHER]`
  \hfill \textit{The poison was placed by an Ashaani agent. The agent wants to destabilise the kingdoms. The water tastes of iron and old regrets. The ninth victim sees the face of the missing king before they die.}
\item \textbf{Peacock Throne Assassination Attempt} --- A courtier tries to kill the king. `[POLITICS]` `[COMBAT]`
  \hfill \textit{The assassin is the chamberlain. The chamberlain is a spy for Vajrapur. He has been waiting nine years. His hands do not shake. He has a ninth dagger hidden in his mirror-stick.}
\item \textbf{Silent Court Leak} --- A judge's verdict is revealed before it is read. `[JUSTICE]` `[SOCIAL]`
  \hfill \textit{The leak is a clerk. The clerk is selling secrets to the Bandit King. The verdict is your death. The clerk is sorry. He is not sorry enough. The ninth page of the verdict is blank — reserved for an appeal from the missing kingdom.}
\item \textbf{Royal Genealogy Forged} --- Someone has altered the king's family tree. The claimant is an impostor — or maybe not. `[POLITICS]`
  \hfill \textit{The forger is a Lethai-thora archivist who lost a bet. The bet was with the Hollow. The Hollow always wins. The ninth generation on the tree is written in disappearing ink. It shows a name that cannot be read.}
\item[J] \textbf{Whispering Pass Ambush} --- The party is caught in a crossfire between two kingdoms. `[COMBAT]` `[STEALTH]`
  \hfill \textit{The pass is neutral ground. Both sides will claim you are a spy. Choose a side — or escape. The wind will carry your choice to both. The ninth whisper in the pass says, “You should have run.”}
\item[Q] \textbf{Immortal Peacock's Cry} --- The peacock screams; the court panics. `[OMEN]` `[FEAR]`
  \hfill \textit{The cry is a warning. An assassin is in the hall. The assassin is a guest. The peacock will not say which one. It wants to watch. The peacock has not screamed in nine years. The last time it screamed, the ninth kingdom fell.}
\item[K] \textbf{Ghost's Demand (Ninth King)} --- The ghost points at you. It accuses you of murdering the king's descendant. `[DEATH]` `[SOCIAL]`
  \hfill \textit{The descendant is an infant. You must prove your innocence. The ghost will not be reasoned with. It will accept a trade: one secret for one life. The ninth secret you offer will be taken by the Hollow.}
\item[A] \textbf{Well's Vision} --- One PC sees their own face in the black water. The face is old, scarred, and weeping. `[MYSTERY]` `[FEAR]`
  \hfill \textit{The well says: “This is your future if you stay.” The vision is true — but only for this moment. The moment is passing. The ninth reflection shows a face that is not yours — it is the missing king's. You feel an obligation you cannot name.}
\end{enumerate}

\section*{Clubs — Highland Complications/Threats}
\label{sec:dhahara-highland-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Cave-In (Sabotage)} --- A tunnel collapses. The collapse was not an accident — a saboteur loosened the timbers. `[DISASTER]`
  \hfill \textit{The saboteur is in the party. Or the party is accused.}
\item \textbf{Toll Dispute} --- The Aeler keeper demands double; the raja's men refuse. The dispute is a trap. `[TOLL]` `[POLITICS]`
  \hfill \textit{Both sides want the party to choose a side. Neither is honest.}
\item \textbf{Ninth Echo} --- On the Bridge of Nine Echoes, someone counts the echoes. The ninth echo says a name. `[OMEN]` `[FEAR]`
  \hfill \textit{The name is the speaker's. They faint. The bridge remembers.}
\item \textbf{Landslide} --- The pass is blocked. The alternative route is three days longer. `[WEATHER]` `[TRAVEL]`
  \hfill \textit{The landslide was deliberate — a rival caravan master wants to delay you.}
\item \textbf{Breath Shortage} --- At high altitude, the party struggles. Mark Fatigue 1 per scene. `[BREATH]` `[ENDURANCE]`
  \hfill \textit{Acclimatisation requires a rest day. The rest day will cost a toll.}
\item \textbf{Spy in Camp} --- A worker is stealing iron samples. The spy is working for an Ashaani agent. `[POLITICS]` `[STEALTH]`
  \hfill \textit{The samples are for a new alloy. The alloy is a weapon.}
\item \textbf{Covenant Audit} --- Sister Gaura demands to inspect the party's cargo. `[LEASE]` `[JUSTICE]`
  \hfill \textit{The cargo includes a relic from the Red Desert — a stolen Covenant seal.}
\item \textbf{Aeler Bell Code Change} --- The toll-keeper rings a different sequence; the party is now ``unregistered.'' `[TOLL]` `[AUTHORITY]`
  \hfill \textit{The change was ordered by the viceroy. The viceroy is testing your loyalty.}
\item \textbf{Elephant Sickness (Curse)} --- A disease spreads through the caravan. The disease is a curse from a local spirit. `[COMMUNITY]` `[MAGIC]`
  \hfill \textit{The spirit was disturbed by the mine's expansion. It wants a sacrifice — a memory, not blood.}
\item[J] \textbf{Viceroy's Summons} --- Thrain demands the party attend him. He offers a deal: find the saboteur in exchange for a toll waiver. `[LEASE]` `[POLITICS]`
  \hfill \textit{The saboteur is his own clerk. The clerk is his daughter.}
\item[Q] \textbf{Silent Pass's Cold} --- The temperature drops; the party loses 1 Supply per hour. `[WEATHER]` `[BREATH]`
  \hfill \textit{The cold is supernatural — a frost spirit has taken residence. It wants a memory.}
\item[K] \textbf{Demon Wakes (Ninth Seal)} --- The sealed mine cracks open. A shadow escapes; it possesses the nearest Aeler engineer. `[MAGIC]` `[LEASE]`
  \hfill \textit{The engineer begins to sabotage the pass. The demon wants the lease broken.}
\item[A] \textbf{Lease Stone Cracks} --- The First Lease Stone glows red and splits. The treaty is void. War is coming. `[LEASE]` `[APOCALYPSE]`
  \hfill \textit{The party has one day to flee the passes before the Aeler and Dhaharan armies clash.}
\end{enumerate}

\paragraph*{(Lowland Rewards)} Peacock feather (enter any court); iron gate pass (brass key); buffalo horn trumpet; Covenant seal (Kosala); temple dancer's anklet (silver); Rajagriha coin; well water (vial); Silent Court disappearing ink (bottle); Nine Kings' key word (“forgiveness”); elephant whistle (bone); immortal peacock's second feather; Whispering Pass's map (ear); Peacock Throne's blessing (decree); ninth king's crown (gold circlet).

\paragraph*{(Highland Rewards)} Aeler breath-counter (silver bell); iron ingot (viceroy's mark); Covenant seal (clay, Himadri); toll waiver (brass nail); elephant whistle (bone); raja's signet ring (silver); smelter's gloves (asbestos); map of the Silent Pass (leather); viceroy's letter of introduction; blind pilgrim's staff; demon's nail (sealed mine timber); ninth echo's whisper (recording); viceroy's forgiveness (pardon).

\section*{Diamonds — Lowland Rewards/Leverage}
\label{sec:dhahara-lowland-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peacock Feather} --- A single iridescent feather from the immortal peacock. Present it to enter any court in the nine kingdoms. `[SOCIAL]`
  \hfill \textit{The feather never fades. It glows when the king is lying. It glows brighter when you are lying. The king will notice. The ninth feather in the peacock's tail is invisible. If you find it, you can command the peacock to speak.}
\item \textbf{Iron Gate Pass (Brass Key)} --- A brass key. Bypass customs at Vajrapur's market. `[TRAVEL]`
  \hfill \textit{The key is also a weapon. It can break chains. It was forged from a slave collar. The slave's ghost thanks you every time you use it. The ghost was a king of the missing ninth. He does not remember his name.}
\item \textbf{Buffalo Horn Trumpet} --- A carved horn. Blow it to summon a buffalo cavalry. `[COMBAT]`
  \hfill \textit{The cavalry are mercenaries. They demand payment in land. One acre per horn blast. They will accept a story instead — if the story is about a lost battlefield. The ninth cavalryman is a ghost. He fights for free, but he will ask you a question after the battle. Answer poorly, and he takes a memory.}
\item \textbf{Covenant Seal (Kosala)} --- A clay seal. Witness any contract in the sacred city. `[JUSTICE]`
  \hfill \textit{The seal is fragile. It breaks if the contract is false. The breaking is loud. The sound carries for a mile. Everyone will know. The ninth seal ever made is still intact. It seals a contract with the missing kingdom. No one has dared to break it.}
\item \textbf{Temple Dancer's Anklet (Silver)} --- A silver anklet with bells. Wear it to gain +1 die to any persuasion roll. `[SOCIAL]`
  \hfill \textit{The bells are silent unless you are honest. Dishonest words make no sound. The silence is more damning than a lie. The ninth bell on the anklet rings only when the monsoon is about to fail. It has rung twice in living memory.}
\item \textbf{Rajagriha Coin} --- A Gupta-era gold coin. Trade it for a favour from the ghosts. `[DEATH]`
  \hfill \textit{The ghosts will answer one question about the empire's fall. The answer is always a warning. The warning is always about you. The ninth coin in the hoard is cursed. Spend it, and a ghost of the missing kingdom will follow you until you return it.}
\item \textbf{Well Water (Vial)} --- A vial of black water from the Nine Kings' Well. Drink it to see the future. `[MYSTERY]`
  \hfill \textit{The future is always a death — yours or someone else's. The vision lasts one minute. You will not forget it. The ninth vial was never drawn. It is reserved for the king who will unite the nine. The well knows your name. It is waiting.}
\item \textbf{Silent Court Disappearing Ink (Bottle)} --- A small bottle. Write a message that vanishes. `[STEALTH]`
  \hfill \textit{The reagent is your own tears. The message reappears when you weep. The reappearance is permanent. Choose your tears carefully. The ninth bottle contains ink that never disappears. It is used to write the names of those who owe the Hollow.}
\item \textbf{Nine Kings' Key Word} --- A single word: “forgiveness.” Speak it to open the tomb. `[LEGEND]`
  \hfill \textit{The word also opens a lock in the Peacock Throne Hall. The lock is a secret door. The door leads to a room that contains the king's true genealogy. The genealogy shows that every king of Ahimsa is a descendant of the missing ninth kingdom. The weight is never lifted.}
\item \textbf{Elephant Whistle (Bone)} --- Carved from a tusk of a battle elephant. Blow it to calm any elephant within earshot — once. `[BEAST]`
  \hfill \textit{The whistle was owned by a mahout who never lost. He is still alive. He wants it back. The elephant that carried the ninth king used a whistle like this. Its ghost can be summoned by blowing it three times at midnight.}
\item[J] \textbf{Immortal Peacock's Second Feather} --- A tail feather from the immortal peacock. It is said to grant a wish. `[LEGEND]` `[MAGIC]`
  \hfill \textit{The wish is that the king will die. Use it wisely. The feather burns after use. The peacock will know. It will not forgive. The ninth wish ever made with a peacock feather was: “Let the missing kingdom return.” The peacock is still deciding.}
\item[Q] \textbf{Whispering Pass's Map (Ear)} --- A chart of the defile and its listening posts. Written on a human ear. `[SPY]` `[TRAVEL]`
  \hfill \textit{The ear was from a spy. It still hears whispers. Hold it to your own ear to listen to distant conversations. The ear will whisper back. Do not answer. The ninth whisper you hear will be your own voice, asking a question you have never asked.}
\item[K] \textbf{Peacock Throne's Blessing (Decree)} --- A royal decree. Demand a favour from any king of the nine. `[AUTHORITY]`
  \hfill \textit{The favour is limited: they will not start a war for you, but they will end one. They will also end a blood feud, a trade embargo, or a marriage alliance. The ninth favour was claimed by the missing kingdom. The king who granted it is still trying to fulfil it.}
\item[A] \textbf{Ninth King's Crown (Gold Circlet)} --- Found in the tomb. Wear it to command the ghosts of Rajagriha. `[DEATH]` `[LEGEND]`
  \hfill \textit{They will fight for you once. Then they will claim you as their king. You cannot leave the ruins. The ruins will grow around you. The ghosts will keep you company. The ninth ghost is the missing king himself. He will ask you one question: “Why did you let me fall?” You have one sentence to answer.}
\end{enumerate}

\section*{Diamonds — Highland Rewards/Leverage}
\label{sec:dhahara-highland-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Aeler Breath-Counter (Silver Bell)} --- Wear it to count your breaths in high altitude. The bell rings when you reach eight. `[BREATH]` `[TRAVEL]`
  \hfill \textit{Stop at eight. The ninth breath is the mountain's.}
\item \textbf{Iron Ingot (Viceroy's Mark)} --- A single bar of high-grade iron, stamped with a dragon. Worth 1 Measure. `[TRADE]` `[LEVERAGE]`
  \hfill \textit{The dragon is the viceroy's mark. Present it to any Aeler forge for a favour.}
\item \textbf{Covenant Seal (Clay, Himadri)} --- A lump of clay. Heat it to witness any oath in the foothills. `[LEASE]` `[JUSTICE]`
  \hfill \textit{The heat comes from a coal taken from the sealed mine. The coal whispers.}
\item \textbf{Toll Waiver (Brass Nail)} --- A brass nail. Present it to pass any Aeler bridge for free. `[TOLL]` `[TRAVEL]`
  \hfill \textit{The nail fits into a hole in the toll-keeper's scale. The hole is for bribes.}
\item \textbf{Elephant Whistle (Bone)} --- A bone whistle. Blow it to call an elephant from the caravan. `[COMMUNITY]` `[TRAVEL]`
  \hfill \textit{The whistle is made from the tusk of a bull that died in a rockslide. Its ghost runs with the elephant.}
\item \textbf{Raja's Signet Ring (Silver)} --- A silver ring with a tiger. Commands loyalty from the raja's clan. `[COMMUNITY]` `[SOCIAL]`
  \hfill \textit{The ring is a copy. The real one was lost in a landslide. The ghost of the original owner wears it.}
\item \textbf{Smelter's Gloves (Asbestos)} --- Handle hot metal safely. The gloves are cursed: they itch when you lie. `[CRAFT]` `[TRUTH]`
  \hfill \textit{The itch is mild at first. It becomes unbearable with repeated lies.}
\item \textbf{Map of the Silent Pass (Leather)} --- A chart drawn on leather. Shows a safe route through the high pass. `[TRAVEL]` `[SECRET]`
  \hfill \textit{The map is incomplete. The missing section is a fold. The fold is a secret — a cave with a spring.}
\item \textbf{Viceroy's Letter of Introduction} --- A sealed scroll. Grants access to any Aeler hold in the mountains. `[LEASE]` `[AUTHORITY]`
  \hfill \textit{The letter is written in blood. The blood is the viceroy's. He changes his blood type each season.}
\item[J] \textbf{Blind Pilgrim's Staff} --- A wooden staff with a bell on top. Ring it to summon a guide. `[FAITH]` `[TRAVEL]`
  \hfill \textit{The guide is a ghost — the pilgrim's predecessor. She will lead you safely, but she will ask for a story.}
\item[Q] \textbf{Demon's Nail (Sealed Mine Timber)} --- A piece of the sealed mine's timber. Touch it to a lock; the lock opens. `[MAGIC]` `[STEALTH]`
  \hfill \textit{The lock also screams. The scream alerts the demon.}
\item[K] \textbf{Ninth Echo's Whisper (Recording)} --- A piece of slate carved with the ninth echo's words. Play it to silence a foe. `[COMBAT]` `[FEAR]`
  \hfill \textit{The foe hears their own name and freezes for one exchange. The slate then crumbles.}
\item[A] \textbf{Viceroy's Forgiveness (Pardon)} --- A written pardon for any crime in the Satrapies. `[LEASE]` `[AUTHORITY]`
  \hfill \textit{The pardon is conditional. The condition is a secret mission. The mission is to assassinate a rival viceroy.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:dhahara-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2–5 → 4, 6–10 → 6, J/Q/K → 8, A → 10).
\item Diamonds are codified outcomes (feathers, keys, seals, bells, ingots) that change position rather than call for a roll.
\item If any A appears, echo \textbf{Dhaharan omens}—peacock feathers shed without reason, a well that reflects a face not yours, a throne that creaks when empty, a whisper that counts to nine, iron that weeps rust, a pass that closes behind you, a bell that rings without a clapper.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:dhahara-mechanics}

\begin{itemize}
\item \textbf{The Nine Kingdoms Pact Clock [8]:} Tracks the cohesion of the lowland alliance. Advances when a kingdom sends a false genealogy, when a king refuses to attend the renewal, or when a border raid occurs. At 4 segments: tariffs increase between kingdoms. At 8: the pact collapses — war is declared, and the Peacock Throne becomes a target. If the Pact Clock ever fills completely, the missing kingdom's ghost appears at the next renewal to cast a vote. No one knows which way it will vote.

\item \textbf{The Ninth Kingdom Clock [8]:} Tracks the return of the missing kingdom. This clock is hidden from players at first, but they may learn of it through research or the Well of Nine Kings. The clock ticks when:
  \begin{itemize}
    \item A king dies without a clear heir (+1)
    \item A genealogy is proven false (+1)
    \item A temple dancer stumbles during the monsoon ritual (+2)
    \item The immortal peacock screams (+2)
    \item A character speaks the name of the missing kingdom aloud (+1, resets monthly)
  \end{itemize}
  At 4 segments: The ghost of the missing king begins appearing in dreams. At 8 segments: The missing kingdom manifests as a phantom realm for one night. The party may enter it — or be claimed by it. Resetting the clock requires a ritual at the Nine Kings' Tomb, paying a weight in memory or blood.

\item \textbf{The Ninth Breath (Highland):} Aeler miners believe that counting to nine in a mine summons the mountain's attention. They count in eights, resetting after each set. Visitors who count to nine aloud may trigger a cave-in — or worse. Mechanically, any character who voluntarily counts to nine in a mine must test Resolve (DV 4) or mark 1 Fatigue as the mountain takes a breath. On a critical failure, the mountain takes a memory instead (lose 1 XP or forget a minor detail). This rule also applies in the lowlands when entering a stepwell or the Nine Kings' Tomb.

\item \textbf{Lease Clock [8] (Highland):} Tracks the political pressure around the treaty renewal. The clock ticks when the party assists or hinders a faction (Aeler, rajas, Admiralty, Covenant). At 4 segments: a faction threatens to walk out of negotiations. At 8: the lease is broken or renewed in the party's favour (or against it). The party can reset the clock by brokering a concession.

\item \textbf{Ancestral Honour (Reputation Track):} Each PC has an Ancestral Honour score from –3 (Heavy Burden) to +3 (Bearer of Light). Gain Burden (move toward negative) by: breaking a sworn oath, failing to protect a family member, disrespecting a temple. Gain Honour (move toward positive) by: keeping a difficult promise, restoring a forgotten genealogy, donating to a royal archive. Effects:
  \begin{itemize}
    \item \textbf{+2 or +3 (Bearer of Light):} +1 Position in social encounters with nobles and priests. You may call in one minor family favour per session.
    \item \textbf{–2 or –3 (Heavy Burden):} Market prices double for you. You may be challenged to a honour-duel (combat or legal contest). If you refuse, you lose 1 permanent Reputation tier.
  \end{itemize}

\item \textbf{Elephant Bonding (Lowland and Highland):} A character who spends a downtime scene with a war elephant may form a bond. The elephant gains a name and a personality. Once bonded, the elephant grants +1 die to the mahout's Command rolls in battle, but the mahout suffers –1 die to social rolls with elephant‑hating factions (e.g., the Aeler counting houses, who fear the beasts). An elephant bonded to a character with high Ancestral Burden (negative) becomes melancholic and refuses to trample enemies of the missing kingdom.

\item \textbf{Ritual Calendar (Lowland):} Temple dancers control the monsoon schedule. A party that earns the favour of a temple dancer (by protecting a shrine, funding a festival, or telling a true story) may ask for a “weather window”: one journey leg where rain, fog, or heat does not apply. Using this favour ticks Ancestral Honour by –1 (the dancer's rivals will spread rumours that you owe a favour to the dancer). If the Ninth Kingdom Clock is at 4+, the weather window may instead open a path to the phantom kingdom.

\item \textbf{Genealogy Forgery (Lowland):} A skilled forger can alter a king's family tree. This requires a 4‑segment Forgery Clock (Wits + Craft, DV 4). If successful, the target king loses 2 Ancestral Honour (becomes more burdened) and the forger gains a permanent String against him. If discovered, the forger is exiled — or executed. A forgery that inserts a connection to the missing kingdom causes the Ninth Kingdom Clock to tick +2 immediately.

\item \textbf{Elephant Community Share (Highland):} Instead of individual ownership, elephant caravans operate on a share system. A character who helps the caravan (clearing a pass, feeding a sick elephant, healing the mahout) gains a \textbf{Share Token}. Spend it to call on the caravan for a single favour: passage for a wagon, a day's fodder, or a warning about a rival. Share Tokens are not weight — they are reciprocity.

\item \textbf{Breath-Bonds as Leverage (Highland):} Aeler engineers use breath-bonds to secure contracts. A character who holds a breath-bond can demand a renegotiation of tolls, quotas, or labour terms. The bond is not a weight — it is a recorded agreement that both parties benefit from keeping. Breaking a breath-bond does not incur a curse; it simply ends the trust, and the other party will never deal with you again.

\item \textbf{The Ninth Caste (Nameless Spies, Lowland):} The ninth caste has no mark, no name, no legal existence. They are the king's enforcers, operating outside the law. A character who discovers a ninth caste spy may leverage that knowledge: gain +1 die on social rolls against that spy's master, but the spy becomes a recurring enemy. The ninth caste has its own obligation to the Hollow; they cannot die until it is paid.

\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Peacock's Eye} — The Witness. *The ruby that sees all false genealogies. It does not blink. It knows who owes.*
    \item \textbf{The Iron Queen} — Mykkiel. *The law that binds the nine kingdoms, even when kings forget. Especially when they forget.*
    \item \textbf{The Monsoon Lord} — Khemesh. *The pressure that brings the rain — and the flood that follows. The flood claims its due.*
    \item \textbf{The Dancing Star} — The Traveler. *The rhythm that guides the dancer's feet and the monsoon's arrival. And the missing step.*
    \item \textbf{The Ninth Throne} — The Hollow. *The empty chair. The unpaid weight. The kingdom that never was.*
    \item \textbf{Thrain the Counter} — Mykkiel (Highland). *The divine auditor; every deed must balance.*
    \item \textbf{The Silent Bell} — The Sealed Gate (Highland). *The threshold that asks only that you stop at eight.*
    \item \textbf{The Forge Father} — Adur (Oath of Flame and Light). *The flame that tempers iron and burns lies.*
    \item \textbf{The Old Breath} — The Traveler (Highland). *The wind that carries caravans and counts their steps.*
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Nine Kings' Tomb and the Kalrantha Amulet (Dhahara):} The tomb is said to contain a map to the Himadri temple where the amulet was sealed. A rival king may offer the party a noble title in exchange for the map. The map is tattooed on the inside of the Ninth King's ghost. You will need to ask politely. The ghost charges a memory for each word of the map.
  \item \textbf{Covenant Judges and the Sidhi (Way of Silk):} High Judge Vasu has ruled on several cases involving Sidhi freedmen. He is known to be fair but harsh. A Sidhi port-master may ask the party to deliver a bribe — or a warning. The warning is written in disappearing ink. The ink is your own blood. The ninth word of the warning is the name of a Sidhi ship that carries cargo for the missing kingdom.
  \item \textbf{Elephant Herds and the Kuvani (Way of Silk):} The Kuvani have tried to buy war elephants from the Nine Kingdoms. The price is always refused — but a Kuvani remount master may offer a story worth the trade. The story must be about a horse that outran a king. The elephants will listen. They are not impressed. The ninth elephant in the herd is a ghost. It remembers the missing kingdom. It will carry anyone who tells it a true regret.
  \item \textbf{The Immortal Peacock and the Hollow (Eastern Realms):} The peacock is said to be the Hollow's original form — before it learned to hunger. Aelaerem hedge-keepers leave offerings of peacock feathers at the Hollow's gates to remind it of what it once was. The feathers are always accepted. The Hollow never thanks them. The ninth feather offered each year is returned, bloodied, to the Peacock Gardens.
  \item \textbf{The Ninth Seal Mine and the Kalrantha Amulet (Dhahara):} The demon sealed in the mine is a fragment of Varash the Unmaker. The Veiled Sisters want to free it to use against the Covenant. The Breath-less want to re-seal it. The demon speaks in the voices of dead miners.
  \item \textbf{Iron Trade and the Ngomebe (Stone Men):} The iron from the Satrapies is used to forge rollers for the walking cities of Ngomebe. A disruption in the iron supply could stop a city mid-migration. The Mason-Kin have a permanent envoy at the Viceroy's hall.
  \item \textbf{The Blind Pilgrim and the Hollow (Eastern Realms):} The pilgrim is searching for a shrine that marks the Hollow's boundary in the high passes. Aelaerem hedge-keepers know the location but refuse to speak it. They say the shrine is a missing ninth step in the mountain's own counting song.
\end{itemize}

\subsection*{Dhaharan Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Peacock:} Crumble a crust before the throne. Ignore the first \emph{Elephant Stampede} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Anklet:} Tie a silver bell to a dancer's anklet. Cancel \emph{Temple Dance Cursed} or take –1 die on your next performance roll (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Well:} Whisper a dead king's name into the black water. Gain +1 Effect when negotiating a truce this scene, but mark \emph{Obligation} to the Covenant. `[SUPERSTITION]`
  \item \textbf{The Ninth Step (Lowland):} When entering a court, take only eight steps. The ninth step is for the ghost of the missing king. If you take it, you will hear a whisper asking for a favour. `[SUPERSTITION]`
  \item \textbf{Bread to the Mine (Highland):} Crumble a crust into the tunnel mouth. Ignore the first \emph{Cave-In} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Bell Rope (Highland):} Tie a knot in the rope of a shrine bell. Cancel \emph{Ninth Echo} or take –1 die on your next stealth roll (the bell still rings, but muffled). `[SUPERSTITION]`
  \item \textbf{Name to the Iron (Highland):} Whisper a dead miner's name into an ingot. Gain +1 Effect when negotiating with Aeler this scene, but mark \emph{Obligation} to the Viceroy. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:dhahara-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{court chamberlains, temple priests, Aeler engineers, or raja's spies}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{elephant mahouts, bazaar merchants, mine guards, or toll-keepers}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{rival spies, Covenant bailiffs, or mountain spirits}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favourable audience with a king or viceroy}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}.
\item \textbf{Ninth Echo (Highland):} A whisper in the pass counts to eight. You feel a presence. The GM gains 1 SB to use on a future supernatural complication.
\item \textbf{Ninth Step (Lowland):} A courtier counts your steps aloud. You stumble. The GM gains 1 SB to use on a future social complication.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local king's guard, Covenant bailiff, mine-captain, or toll-keeper} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{elephant stampede, riot, landslide, or thin air}.
\item \textbf{Foe Appears:} A \textbf{rival kingdom's spy, bandit, or demon fragment} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\item \textbf{Genealogy Taint (Lowland):} A rumour spreads that you are descended from the missing kingdom. Social rolls with nobles are Desperate for one scene.
\item \textbf{Breath Shortage (Highland):} The air thins. All characters mark 1 Fatigue.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{elephant cavalry, temple guards, bandit crews, Aeler engineers, or raja's soldiers} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes—\textbf{pact broken/well poisoned/ghost accuses/mine seals/lease stone cracks/ninth echo spoken}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\item \textbf{Ninth Kingdom Clock ticks +1:} The missing kingdom stirs. A ghost is seen. A temple dancer stumbles. A miner counts to nine.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Maharaja, High Judge Vasu, Ninth King's Ghost, Viceroy, or Demon from the Ninth Seal} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically—\textbf{throne hall collapses/well floods/peacock screams/mine collapses/bridge falls/lease stone shatters}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice—\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\item \textbf{The Missing Kingdom Manifests:} For one scene, the phantom realm overlaps with reality. All characters must test Resolve (DV 5) or become Lost (cannot be found until dawn).
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Dhahara (Court Law):} Peacock feathers shed without reason, a throne's ruby eyes follow a different person, a forgotten genealogy appears on a wall.
\item \textbf{Dhahara (Elephant Law):} Elephants kneel unprompted, a tusk glows faintly, a mahout speaks a prophecy in his sleep.
\item \textbf{Dhahara (Ancestral Honour):} A merchant refuses to meet your eyes, a child throws salt at your feet, a temple door closes as you approach, a ghost asks when you will pay what you owe.
\item \textbf{Dhahara (Mine Law):} Iron veins glow faintly, the mountain groans, a breath-counter ticks backward.
\item \textbf{Dhahara (Lease Politics):} The Viceroy's bell rings once for no reason, a raja's elephant kneels unprompted, a treaty scroll spontaneously combusts.
\item \textbf{Dhahara (Pass Law):} Snow falls upward, the ninth echo repeats a different name each time, a bridge appears where there was only chasm.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth or Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Dust Cloud} (lowland) or \emph{Thin Air} (highland).
\item \textbf{Cut the Pact (Presence + Sway):} On a hit, freeze range for one exchange; if a peacock feather is in your possession, +1 die.
\item \textbf{Cut the Toll (Presence + Sway, Highland):} On a hit, freeze range for one exchange; if a toll waiver is in your possession, +1 die.
\item \textbf{Weather the Stampede (Body + Ride):} On a hit in \emph{Elephant Stampede}, you pick who is trampled; on a miss, the GM does.
\item \textbf{Weather the Cave-in (Body + Athletics, Highland):} On a hit in \emph{Cave-In}, you pick who escapes first; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Throne, Well, Mine, and the Missing Ninth)}
\begin{itemize}
\item \textbf{The Ninth King (Unburied)} — \emph{Phases}: Tomb Robberies $\rightarrow$ Orchestrated Succession Crisis $\rightarrow$ Throne Coup. \emph{Cracks}: return a stolen crown, expose a false genealogy, survive the Ninth King's Ghost. \emph{Secret}: The Ninth King does not want to be restored to power — he wants his obligation to the Hollow paid. Pay it, and he will rest.

\item \textbf{The Drought-Maker (Well Poisoner)} — \emph{Phases}: Wells Drying $\rightarrow$ Procession of Empty Stepwells $\rightarrow$ Monsoon Baptism of a Dancer. \emph{Cracks}: vial of pure well water, survivor testimony, the temple dancer's bells ringing true again. \emph{Twist}: The Drought-Maker is a servant of the missing kingdom, sent to weaken the eight so that the ninth may return.

\item \textbf{The Ghost Emperor of Rajagriha} — \emph{Phases}: Imperial Echo $\rightarrow$ Legion of Shades $\rightarrow$ The Unfinished Throne. \emph{Cracks}: speak a true kindness about his empire, return a broken pillar to its original place, promise to remember a name he has forgotten. \emph{Price}: He will not pass on until someone takes his place as keeper of the missing kingdom's memory.

\item \textbf{The Hollow's Accountant (Ninth Caste Master)} — \emph{Phases}: Audit of the Living $\rightarrow$ Repossession $\rightarrow$ Foreclosure on a Kingdom. \emph{Cracks}: prove that a weight has been paid twice, present a genealogy that shows the Hollow itself as a debtor, or offer a new missing kingdom to take the place of the old. \emph{Warning}: The Accountant cannot be fought — only negotiated with. Its quill writes in reverse. Its ink is your blood.

\item \textbf{The Breath-Eater (Nine Silence, Highland)} — \emph{Phases}: Mines Collapsing $\rightarrow$ Orchestrated Cave-In $\rightarrow$ Pass Coup. \emph{Cracks}: find a lost breath-counter, expose a false lease, survive a count of nine.

\item \textbf{The Unsealed (Demon of the Ninth Seal, Highland)} — \emph{Phases}: Whispers in the Mine $\rightarrow$ Procession of Cracking Timbers $\rightarrow$ Iron-Baptism of a Viceroy. \emph{Cracks}: relic Covenant seal, survivor testimony, the mine's face weeping clear water.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Dhaharan, power flows through genealogy, elephant herds, the honour of the court, and the crushing weight of ancestral obligation — and in the high passes, through iron, tea, and the lease. A king's authority rests on his dynasty's archive; a queen's on her marriage alliances; a temple dancer's on the rhythm of the monsoon; a viceroy's on the High King's seal and the loyalty of his engineers. The Peacock Throne is the centre, but the other eight kingdoms are not satellites — they are competitors. And the missing ninth is the weight that never stops compounding. In the north, the Aeler and Dhaharan have bled into the same stone, and the mountain remembers both.

\textbf{For the GM:}  
Power in Dhahara is measured in iron, buffalo, the thickness of a family's scroll collection, the number of forgotten promises in its bloodline, and the number of breaths a miner can hold. Patronage often takes the form of peacock feathers, iron gate passes, Covenant seals, royal decrees, breath-counters, and toll waivers — each tied to a specific kingdom, clan, court, or mine. To emphasise the honour-based strangeness and the uneasy Aeler-Dhaharan hybrid:

\begin{itemize}
\item Use the \textbf{Ancestral Honour Dial} as a visible track. Let players gain or lose Honour based on their actions in native courts. High Honour (positive) opens doors; low Honour (negative) closes them — and may bring the Hollow's accountants.
\item Let the nine kingdoms be a web of shifting alliances and unpaid obligations. A favour from the Maharaja may be an insult to the Rani — and a weight to a third kingdom that must be balanced.
\item Use the stepwells, elephant stables, throne halls, mines, bridges, and the Viceroy's hall as arenas for social contests, where a well-recited genealogy is a weapon, a broken promise is a declaration of war, and a count of eight is a threat.
\item Track the \textbf{Ninth Kingdom Clock} as a slow-burn campaign timer. The party cannot stop it permanently — only delay it, redirect it, or prepare for the night when the missing kingdom returns. That night, the Hollow will collect. The question is who will pay.
\item In the high passes, use the \textbf{Lease Clock} to mirror the politics of the lowlands. The renewal of the lease is a recurring pressure point; the party can tilt it toward Aeler, rajas, or a third party.
\item Emphasise that the Aeler are \textbf{not conquerors} — they are lessees, bound by treaty. Their engineers and viceroys are in Dhaharan territory on Dhaharan sufferance, renewed every nine years. The rajas tolerate them because the iron trade is profitable and the passes are dangerous. But the lease can be broken — by a stone cracking, by a demon waking, by a king's whim.
\end{itemize}
In Dhahara, the peacock remembers every eye, the well shows every death, the mountain takes every ninth breath, and the missing kingdom waits. Its weight is never closed.
\end{tcolorbox}

% Index entries
\index{Dhahara|see{Elephant Kingdoms, Mountain Satrapies, Nine Kingdoms}}
\index{Theme generator}
\index{Dhaharan generator}
\index{Mountain Satrapies}
\index{Nine Kingdoms}
\index{Peacock Throne}
\index{Elephant warfare}
\index{Covenant judges}
\index{Ancestral Honour}
\index{Ninth Kingdom Clock}
\index{Lease Clock}
\index{Ninth Breath}
\index{Elephant Bonding}
\index{Local Patrons!Dhahara}