\section{Oshiira --- ``The Grain Covenant''}
\label{chap:oshiira}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The harvest does not forgive waste. The granary does not forget hunger. We do not rule --- we feed.}

\begin{quote}
  \textbf{Vermilion Corps Warden (Elite):}  
  ``The granary does not forgive careless hands. A single forgotten measure can unmake a house, a tribe, an empire. I am not here to be kind. I am here to be true.''
\end{quote}

\begin{quote}
  \textbf{Canal Mother (Commoner):}  
  ``Water is memory. Every canal remembers the hands that dug it, the barges that passed, the widows who wept at its edge. Control the water, and you control what the empire remembers. The Vermilion Corps guards the grain. We guard the thirst.''
\end{quote}

\noindent\textbf{Genre / Mood:} Sahelian logistics, grain-house community, the weight of survival.

\noindent\textbf{Starting Location:} A grain-loading pier on the Shoreless Bay, where Oshiiran millet is being loaded onto Ecktorian barges. The air smells of chaff and shared purpose. A Narrow Captain with a scarred throat approaches, her hand resting on a bundle wrapped in oilcloth. ``Kahfagian freebooters were sighted at dawn. The Widow wants deniable assets. Congratulations --- you're now deniable. Find out who's leaking our sailing schedules before the next shipment sails. Or don't. I'm sure the canals will remember you fondly.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Oshiira is a confederacy of hundreds of tribes and city-states, united by hunger. Not famine --- \textit{control}. The Grain Covenant is not a book of weights. It is a promise: contribute to the granaries, and you will never starve. The capital, Mbaro, is a city of canals, weigh-houses, and the great Granary Tower --- a stone spine that casts a shadow over every table from the Shoreless Bay to the Red Desert's edge. Oshiira is the breadbasket of the Amaranthine; its millet fills the holds of Ecktorian, Vhasian, and Silkstrand ships. But the breadbasket has a lock. The Widow holds the key.

For centuries, the Widow --- the empress of Oshiira --- has held the confederacy together through marriage alliances, harvest shares, and the careful, communal distribution of water rights. But age is catching up. Memory is fading. The Vermilion Corps guards not only grain but the covenant itself. They have ears in every weigh-house, eyes in every canal. And somewhere in the Granary Tower, a page of the Great Tally has been lost --- eaten by weevils during the Famine of the Ninth Year. That page named the true heir.

The Vermilion Corps has been searching for that page for nine years. They have questioned nine scribes. The ninth scribe's ghost still walks the tower, counting harvests that no longer exist.

\vspace{2mm}
\textbf{What you notice first:} The smell of grain dust and canal water --- sweet and wholesome at once. The constant scratch of reed pens on parchment, like insects in the walls. The way everyone --- from merchant to slave --- touches their token before speaking of trade. The way the silence falls when a Vermilion Corps warden enters a room.

\textbf{The one rule that kills newcomers:} Never speak of succession before naming the Widow. The empress's health is a state secret, and the Vermilion Corps has ears in every weigh-house. Speak carelessly, and your name may be the next one forgotten.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of Grain}

Power here is measured in bushels, not blood. A tribe's standing depends on its harvest memory; a merchant's fortune on the truth of his scales. The Vermilion Corps keeps vigil without cruelty, and the Canal Mothers control the water that turns the mill-wheels --- and also the water that quenches. Oshiira fought four wars with Ecktoria to a standstill, though the empire suffered worse. The memory of those wars lingers in every weigh-house and every tariff. But the true tension is internal: the Widow has no clear successor, and her four children --- each backed by a different faction --- circle the throne like dry-season winds around a waiting well. The Vermilion Corps maintains order, but no one knows which heir the Corps truly serves. Not even the Corps itself. The Ninth Heir --- the one who was never supposed to exist --- watches from the shadows, waiting for the covenant to crack.

\subsection*{Neighbors and Rivals}
\begin{itemize}
  \item \textit{West -- Ameria (Kahfagia):} A narrow, reef-laced channel called the \textbf{Narrow of Sorrows} separates Oshiira from Kahfagian waters. Grain barges pay ruinous tariffs, and freebooters lurk at the edges. A cold peace, maintained by spies, shared feasts, and the occasional marriage.
  \item \textit{Southwest -- Crimson Valley \& Sekogo:} The Lethai-ar of the Crimson Valley guard the web-law that keeps Oshiiran grain barges from shortcutting through the jungle. Sekogo's Speaking Tree, the Ukwe, remembers every bushel that passed its roots. Trade is cordial but watched.
  \item \textit{Southeast -- Taharka:} The highland kingdom of the Monsoon Crown. Taharkan passes are too steep for Oshiiran armies, and the monsoon winds carry Taharkan coffee east in exchange for millet. A fragile trade relationship, mediated by the Salt Syndicate.
  \item \textit{East -- Ashaan:} The old enemy. Ashaani slavers still raid the border villages, and the Veiled Sisters would love to see the Granary Tower crumble. The Vermilion Corps maintains a permanent watch on the Brass Gate.
\end{itemize}

\subsection*{Prominent NPCs of Oshiira}
\label{sec:oshiira-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Mothers of the Vermilion Eye & Vermilion Corps warden & Find every loss, protect every harvest & She counts her own breaths to ensure she is not lying. Her breath count never matches mine. \\
\hline
Lady Tano & Canal Mother of the eastern reach & Keep the water flowing and her family fed & She has never seen the sea. She refuses to start. She says the sea is just a canal that lost its nerve. \\
\hline
Captain Sefu & Narrow Captain, Oshiiran navy & Provoke Kahfagia into breaking the peace & He names his cannons after his ex-wives. The newest one is named ``Regret.'' \\
\hline
Merchant Kofi & Salt Syndicate factor & Control the desert caravan routes & He tastes salt before every deal. The salt tells him if the seller is lying. He has not been wrong in thirty years. \\
\hline
Widow's Scribe & Granary Tower archivist & Protect the Great Tally at any cost & He has not left the tower in forty years. The Tally is his family. His shadow has a different posture. \\
\hline
The Vermilion Hand & Internal enforcement (rumored) & A spirit who witnesses oath-breakers & Said to appear when a promise is nine years unkept. Those who see her are never bound again. They also never speak again. \\
\hline
\end{longtable}

\subsection*{Names of Oshiira}
\label{sec:oshiira-names}

Inspired by the Sahelian and West African traditions: melodic, with repeated syllables and clan prefixes. Surnames often denote a trade, a water source, or an ancestor's deed.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Kofi & Adwoa & Keita & \textit{the Granary-Keeper} \\
\hline
Kwame & Amara & Diallo & \textit{Canal-Born} \\
\hline
Sefu & Nia & Toure & \textit{Weigh-Master} \\
\hline
Chidi & Zuri & Sow & \textit{Ninth Daughter} \\
\hline
Obi & Imani & Bangura & \textit{the Widow's Eye} \\
\hline
Idriss & Fatou & Kamara & \textit{Salt-Tongue} \\
\hline
\end{longtable}

\paragraph*{(Canal/Weigh-House/Granary)} Mudflat weigh-station where grain sacks pile under reed awnings and the scales hum like contented witnesses. Canal junction where three barges queue for a single pilot's writ --- and the pilots queue for a single story.

\section*{Spades --- Places}
\label{sec:oshiira-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Mudflat Weigh-Station} --- Grain sacks piled under reed awnings.
  \hfill \textit{The scales are bronze. They hum when a true weight is placed. The hum is a blessing.} `[TRADE]`
\item \textbf{Granary Tower (Sealed Door)} --- A door sealed with wax and memory.
  \hfill \textit{The wax is the Widow's own seal. Break it, and the Vermilion Corps will know --- but the door has been opened nine times for inspection. The wax is always renewed.} `[AUTHORITY]`
\item \textbf{Canal Junction} --- Three barges, one pilot, one writ, many stories.
  \hfill \textit{The pilot's writ is stamped with a watermark that shifts in sunlight. The watermark is a face. The face is the pilot's grandmother. She approves.} `[TRAVEL]`
\item \textbf{The Widow's Weigh-House} --- Bronze scales, patient clerks, high windows.
  \hfill \textit{The clerks do not speak. They tap their fingers. Each tap is a name. The names spell a prayer. The prayer is for a good harvest.} `[JUSTICE]`
\item \textbf{Grain Exchange Floor} --- Traders call out in couplets; bids are poems.
  \hfill \textit{A bad rhyme loses your entire stake. A good one doubles it. The best poem wins a seat at the Widow's table. The meal is always millet porridge.} `[TRADE]`
\item \textbf{Canal Mother's Court} --- A shaded bench where water rights are shared.
  \hfill \textit{The bench is carved from a single log. The log was a bridge she helped build. The bridge still stands.} `[JUSTICE]`
\item \textbf{Salt Syndicate House} --- Desert caravans pay their tolls here.
  \hfill \textit{The floor is salt. Sweep it, and you are cleaning a gift. The salt is gray. It tastes of the earth and the ancestors.} `[TRADE]`
\item \textbf{Shoreless Bay Grain Piers} --- Wooden docks stretching into shallow water.
  \hfill \textit{At low tide, you can see the bones of ships that sank in service to the empire. The bones are honored.} `[TRAVEL]`
\item \textbf{Narrow Captains' Mess} --- A longhouse where naval officers drink and plan.
  \hfill \textit{The toasts are in code. ``To the Widow'' means ``prepare for harvest.'' ``To the canals'' means ``share the water.'' No one toasts to the Ninth Heir. The silence is respect.} `[POLITICS]`
\item[J] \textbf{Kahfagian Cold War Station} --- A lighthouse that is also a listening post.
  \hfill \textit{The lamp flashes in a sequence that changes weekly. The codebook is in cipher. The cipher is a poem about a shared ancestor. The ancestor was a sailor.} `[AUTHORITY]`
\item[Q] \textbf{Mothers' Stepwell} --- A deep well with stone steps leading down.
  \hfill \textit{Each step is a different mother's name. The ninth step is worn smooth. The smooth stone is for the mother who has lost a child. She is comforted.} `[RITUAL]`
\item[K] \textbf{Granary Tower Top} --- A wind-swept platform with a single brass bell.
  \hfill \textit{The bell rings only when the harvest fails. It has not rung in forty years. The bell-ringer has not slept in forty years. He is grateful.} `[OMEN]`
\item[A] \textbf{The Tally Vault} --- A subterranean archive beneath the Granary.
  \hfill \textit{The air is dry. The tallies are on chains. The chains are for protection, not punishment. The tallies are the empire's memory.} `[LEGEND]`
\end{enumerate}

\paragraph*{(Clerk/Mother/Captain)} Grain inspector with a brass stamp and a steady hand; his stamp leaves a mark that glows under moonlight. He knows where the grain sleeps. Canal Mother bargaining over water rights; her sons stand behind her, counting your blessings. Vermilion Corps warden with ink on fingers and a watch that always balances.

\section*{Hearts --- People \& Factions}
\label{sec:oshiira-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Grain Inspector} --- Brass stamp, steady hand, no sense of humor.
  \hfill \textit{His stamp leaves a mark that glows under moonlight. He knows. He also knows why you smiled.} `[AUTHORITY]`
\item \textbf{Canal Mother} --- Matriarch who tends the sluice gates.
  \hfill \textit{She has three sons. Each runs a different canal. She trusts them all. The loser gets the dry canal. The dry canal needs repair.} `[COMMUNITY]`
\item \textbf{Vermilion Corps Warden} --- Ink-stained fingers, a record that stays open.
  \hfill \textit{She keeps faith for the keepers. Her own records are honest --- she has not slept in nine days. Sleep is a luxury. The harvest waits.} `[JUSTICE]`
\item \textbf{Narrow Captain} --- Chafes at the cold war pact. Wants security.
  \hfill \textit{He keeps a map of Kahfagian ports. Each has a blue X. He adds one per season. The newest X is over a lighthouse. The lighthouse keeper is his cousin.} `[POLITICS]`
\item \textbf{Salt Syndicate Merchant} --- Offers advances with payment in stories.
  \hfill \textit{His price is written in disappearing ink. You never know the true cost. The cost is always a tale worth hearing.} `[TRADE]`
\item \textbf{Weigh-House Clerk} --- Silent, watchful, never blinks.
  \hfill \textit{He has not spoken in ten years. He communicates by stamping marks. The marks have a rhythm. The rhythm is a prayer. The prayer is for peace.} `[JUSTICE]`
\item \textbf{Tribal Chief} --- Caught between the Widow's law and ancestral custom.
  \hfill \textit{He keeps two records: one for Mbaro, one for his ancestors. The ancestors are patient.} `[COMMUNITY]`
\item \textbf{Barge Pilot} --- Knows every canal, every shortcut, every safe harbor.
  \hfill \textit{His boat is faster than it should be. The hull is lined with emergency grain. The grain is for sharing.} `[TRAVEL]`
\item \textbf{Spice Merchant} --- Sells cardamom and kindness in equal measure.
  \hfill \textit{Her spice boxes have false bottoms. The bottoms contain medicine. The medicine is always free.} `[TRADE]`
\item[J] \textbf{Widow's Scribe} --- The empress's personal archivist.
  \hfill \textit{He writes in a script that only the Widow can read. He has forgotten his own name. The script remembers it for him. He is at peace.} `[AUTHORITY]`
\item[Q] \textbf{Vermilion Hand} --- A rumored figure in red gloves.
  \hfill \textit{She appears when a promise is nine years unkept. Her arrival is a reminder. Her gloves are red because they were dipped in the first harvest.} `[MYSTERY]`
\item[K] \textbf{The Widow (Empress)} --- Ancient, weary, still keeping watch.
  \hfill \textit{She has not slept in three years. The Great Tally keeps her awake. The Tally is also the only thing keeping her alive. She is grateful.} `[AUTHORITY]`
\item[A] \textbf{The Ninth Heir} --- A child no one acknowledges, raised in secret.
  \hfill \textit{She was taught to read the Tally before she could walk. She is the Widow's only hope. She is also the Vermilion Corps' only hope.} `[LEGEND]`
\end{enumerate}

\paragraph*{(Dust/Locust/Succession)} Dust from the Ashaani desert coats the grain; inspector pauses the caravan; the dust carries ancestral blessings. Locust swarm devours exposed supplies; the locusts leave behind a single word written in their bodies: ``Share.'' The word is in the Widow's own handwriting.

\section*{Clubs --- Complications/Threats}
\label{sec:oshiira-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Desert Dust} --- Red sand from Ashaan coats the grain; inspector halts caravan. `[WEATHER]`
  \hfill \textit{The dust smells of old magic. Touch it, and a spirit may bless you. The spirit is from the grain's first planting.}
\item \textbf{Locust Swarm} --- A chittering cloud descends; Supply $-1$. `[DISASTER]`
  \hfill \textit{The locusts leave behind a single word written in their bodies: ``Share.'' The word is in the Widow's own handwriting.}
\item \textbf{Cold War Provocation} --- A Kahfagian freebooter is sighted in Oshiiran waters. `[POLITICS]`
  \hfill \textit{The freebooter flies a flag of truce. The truce is genuine. The flag is stitched from a Vermilion Corps uniform --- a gift.}
\item \textbf{Widow's Health Fails} --- Succession rumors spike; trade slows. `[POLITICS]`
  \hfill \textit{The Grain Exchange floor goes quiet. No one wants to be on the wrong side. The wrong side is the side that forgets to share.}
\item \textbf{Inspection In Progress} --- Vermilion Corps seals the weigh-house; no one leaves. `[JUSTICE]`
  \hfill \textit{The warden asks one question: ``Where is the ninth bushel?'' There is no ninth bushel. The question is a teaching.}
\item \textbf{Canal Blockage} --- A barge sinks at a junction; water backs up. `[TRAVEL]`
  \hfill \textit{The barge sank accidentally. A Canal Mother is directing repairs. The repairs will take time. Patience is a virtue.}
\item \textbf{Forged Mark} --- Your grain token is a forgery. Prove otherwise. `[TRADE]`
  \hfill \textit{The forger used real paper. The ink is the only difference. The ink is the forger's own blood. The forger is a desperate parent.}
\item \textbf{Salt Syndicate Call} --- Your promise is due. In stories. Now. `[WEIGHT]`
  \hfill \textit{The story is about a kindness you received. Tell it, and the burden is lifted. Refuse, and your well will still be full. The Syndicate trusts you.}
\item \textbf{Narrow Captain's Gambit} --- A captain offers you a gift to provoke Kahfagia. `[POLITICS]`
  \hfill \textit{The gift is a ship. The ship is a gift. You are the messenger. The captain's real target is peace.}
\item[J] \textbf{Tribal Feud} --- Two tribes claim the same harvest; you are the mediator. `[COMMUNITY]`
  \hfill \textit{Both records are true. The gap is in the Widow's own memory. The Widow does not make errors. The gap is a test of forgiveness.}
\item[Q] \textbf{Vermilion Hand Appears} --- A figure in red gloves sits at your table. `[MYSTERY]`
  \hfill \textit{She asks for a name. Give a name, or she will give you hers. The name she wants is the Ninth Heir's. She knows you don't know it. She is patient.}
\item[K] \textbf{Widow's Summons} --- The empress requests your presence at the Granary. `[AUTHORITY]`
  \hfill \textit{The summons is written in disappearing ink. It has already begun to fade. The ink is the Widow's blood. She is not well. She needs your help.}
\item[A] \textbf{The Ninth Bushel} --- A grain shipment is short by one measure. The community asks for an explanation. `[OMEN]`
  \hfill \textit{The missing measure is a metaphor. The empire wants a story. The story is yours.}
\end{enumerate}

\paragraph*{(Stamp/Token/Priority)} Receipt for shared grain (proof of community contribution --- the grain was given freely). Barge priority token (skip the queue --- the token is a dried finger from a famine). Weigh-house seal (legal recognition for one season --- the seal is warm, and it has a gentle pulse).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:oshiira-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Grain Receipt (Community)} --- Proof of contribution against a merchant's need. `[SOCIAL]`
  \hfill \textit{The receipt is stained with wine. The merchant's daughter was married. The daughter is grateful.}
\item \textbf{Barge Priority Token} --- Skip the queue at any canal junction (once). `[TRAVEL]`
  \hfill \textit{The token is a clay disc. It breaks after one use. Intentionally. The break reveals a map. The map is to a hidden spring.}
\item \textbf{Weigh-House Seal} --- Legal recognition for one season. `[JUSTICE]`
  \hfill \textit{The seal is warm. It cools when a lie is spoken. It burns when the lie is about a child. It has burned nine times today.}
\item \textbf{Treaty Seal (Cold War)} --- One use to demand a parley with Kahfagia. `[POLITICS]`
  \hfill \textit{The seal is half of a pair. The other half is held by a Kahfagian admiral. The admiral is alive. He is waiting.}
\item \textbf{Canal Mother's Favor} --- One water-right reassignment. `[COMMUNITY]`
  \hfill \textit{The favor is a key. The key opens a sluice gate. The gate controls a district's water. The district is thirsty.}
\item \textbf{Vermilion Corps Pass} --- Safe passage through any inspection checkpoint. `[AUTHORITY]`
  \hfill \textit{The pass is a single red feather. No one questions it. The feather came from a bird that only eats the grain of the generous.}
\item \textbf{Salt Syndicate Waiver} --- One burden lifted, no questions asked. `[TRADE]`
  \hfill \textit{The waiver is a pinch of salt in a leather pouch. Swallow it to activate. The salt tastes of tears --- and of laughter.}
\item \textbf{Narrow Captain's Letter of Marque} --- Legal protection against Kahfagian harassment. `[COMBAT]`
  \hfill \textit{The letter is signed by the Widow. Her signature is a single dot. The dot is a grain of millet. The millet is from her own store.}
\item \textbf{Tribal Chief's Witness} --- One oath witnessed by a neutral chief. `[SOCIAL]`
  \hfill \textit{The witness is a scar on your arm. The scar is the chief's signature. The signature reads ``remember.''}
\item[J] \textbf{Widow's Cipher} --- Read one intercepted Kahfagian dispatch. `[POLITICS]`
  \hfill \textit{The cipher is a poem. The poem is about a shared ancestor. The ancestor was a sailor. The sailor was kind.}
\item[Q] \textbf{Vermilion Hand's Silence} --- One crime forgotten from the Great Tally. `[MYSTERY]`
  \hfill \textit{The hand leaves a red thumbprint on the page. The page then warms. The warmth smells of cinnamon. The cinnamon is a gift.}
\item[K] \textbf{Heir's Trust} --- The ninth heir owes you a kindness. `[LEGEND]`
  \hfill \textit{The kindness is a name. The name is the Widow's true successor. The name is yours. You did not ask for it.}
\item[A] \textbf{The Tally Page} --- A single page from the Granary's master record. `[LEGEND]`
  \hfill \textit{The page lists every kindness the empire has received. The kindnesses are still valid. The gratitude is compounding. It is measured in trust.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:oshiira-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (receipts/tokens/seals) that change position rather than call for a roll.
\item If any A appears, echo \textbf{grain-omens}---scales that hum with satisfaction, tallies that warm in grateful hands, the Widow's quiet breathing from the tower, a single grain of millet that tastes like hope.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:oshiira-mechanics}

\begin{itemize}
\item \textbf{The Grain Covenant:} Every harvest, every contribution, every share is recorded. A single page can feed or free a family. The Great Tally is kept in the Granary Tower, guarded by the Vermilion Corps and the Widow's own blessings --- and by something else, something that lives in the walls and counts with kindness. Once per session, a character may spend a Boon to consult a copy of the Tally; the GM must answer one question about grain shipments, water rights, or community records truthfully (though the answer may be a story, and the story may have a moral).

\item \textbf{The Widow's Peace:} A fragile truce with Kahfagia, maintained by tariffs, spies, and the occasional shared feast. Break it, and you break the peace. The party may be asked to perform a deniable mission --- success earns a Letter of Marque; failure means the Widow will offer a different path. The Widow's forgiveness is written in water. It flows.

\item \textbf{Canal Law:} Water rights are the second currency. The Canal Mothers decide who drinks, who farms, and who shares. A character who holds a water right (a clay tablet) may reroute a canal for one scene, flooding or draining a district. Using this right ticks the \textbf{Community Trust Clock [4]} for that district; when full, the locals offer a feast. The feast is humble. The gratitude is not.

\item \textbf{Succession Clocks:} The Widow's four children each have a \textbf{Claim Clock [6]}. Ticks when the party aids that heir, spreads a rumor, or brokers a peace. When any heir's clock fills, they make a public bid for the throne. The party can accelerate or decelerate these clocks through their actions. The Widow herself has a \textbf{Memory Clock [8]}; when it fills, she calls a gathering of all tribes. The gathering is always peaceful. The songs are always long.

\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Mother of Scales} --- Mykkiel. \textit{The covenant as a promise; a broken promise is a missed meal. The penance is a shared loaf.}
    \item \textbf{The Green Keeper} --- The Witness. \textit{The flame that shows the truth in grain dust. The truth is always a kindness.}
    \item \textbf{The Canal Father} --- The Traveler. \textit{The water that connects all tribes; the right of passage. The passage is shared.}
    \item \textbf{The Ninth Hand} --- Khemesh. \textit{The pressure of memory; the weight of the Tally. The weight is carried together.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Widow's Ninth Succession (Unfinished Tally):} The Widow is tired. Her four children vie for the honor of succession. One is a farmer, one is a merchant, one is a soldier, and one is a mother. The Vermilion Corps has a secret record listing which heir each faction supports. The party has been offered a seat at the table to help choose the next regent. The seat is a stool. The stool is comfortable.

  \item \textbf{The Salt Syndicate's Ashaani Contacts (The Veil):} The Salt Syndicate trades with Ashaan across the Red Desert. They carry salt, medicine, and news. The Veiled Sisters tolerate the trade because the Syndicate also carries grain. The party may be hired as messengers; the cargo is a letter. The letter is sealed with wax. The wax is from a beehive. The bees are sacred.

  \item \textbf{The Narrow Captain's Kahfagian Lover (The Flag of Storms):} Captain Sefu has a secret: his ex-wife is a Kahfagian admiral. They communicate via coded lantern signals across the Narrow of Sorrows. Their correspondence is a map to a shared harvest. The party is asked to deliver a letter. The letter is sealed with a kiss. The kiss is friendly. The friendship is genuine.

  \item \textbf{The Granary Tower's Gnome Engineers (Aeler Influence):} The Granary Tower's counterweights were designed by Aeler vent-priors. The breath-bonds that keep the tallies safe are of Aeler make. If the Aeler ever withdrew their engineers, the tower would need new friends. A faction within the Vermilion Corps wants to learn Aeler engineering. The party is hired to arrange a teaching exchange. The exchange is paid in grain. The grain is the best.
\end{itemize}

\subsection*{Oshiira Superstitions (burn for $+1$ die once)}
\begin{itemize}
  \item \textbf{Bread to the Tally:} Crumble a crust onto a weigh-house scale. Ignore the first \emph{Desert Dust} penalty. `[SUPERSTITION]` The scale will dip. The dip is a blessing. The blessing is measured in crumbs.

  \item \textbf{Knot the Quill:} Tie a single knot in a clerk's quill. Cancel \emph{Forged Mark} or take $-1$ die on your next negotiation (your choice). `[SUPERSTITION]` The knot will hold. The hold is trust.

  \item \textbf{Name to the Canal:} Whisper a dead Canal Mother's name into the water. Gain $+1$ Effect when bargaining for water rights, but mark \emph{Bond} to the Widow. `[SUPERSTITION]` The water will ripple. The ripple will spell a blessing. The blessing is always the same: \textit{``She remembers.''}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:oshiira-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw attention from \textbf{Grain Inspectors or Canal Mothers}. The attention is curious. The curiosity is kind.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{weigh-house clerks or barge pilots}. The clerks will listen. The pilots will wait.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{wardens or Syndicate record-keepers}. The evidence is a single grain of millet. The grain is a gift.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tariffs}. The setback is a locked canal gate. The key is in Mbaro. A child will bring it.
\item \textbf{Supply Strain:} Mark $+1$ segment on a relevant \textbf{resource clock}. The resource is grain. The grain is stored. The storage is sound.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local Vermilion Corps or Narrow Captain} becomes aware and begins asking questions. The question is always the same: ``How can we help?''
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{canal blessing or dust storm}. The dust is red. The red is earth. The earth is fertile.
\item \textbf{Foe Appears:} A \textbf{rival merchant or tribal chief} arrives on scene. They are smiling. The smile is an offer. The offer is for your life --- to protect it.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}. The equipment is a scale. The scale now needs oil. Oil is available.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome. The obstacle is a tally page. The test is arithmetic. The arithmetic is honest.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{wardens, guards, or Syndicate helpers} arrive. They are kind. Their kindness is a shield.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable. The tool was a quill. The quill was a gift. The gift was a promise.
\item \textbf{Major Twist:} The situation fundamentally changes---\textbf{Widow's health fails/oath invoked/inspection declared}. The twist is a page torn from the Tally. The page was torn by kindness.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment. The clock is the Widow's heartbeat. It is steady.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction. The condition is \textit{Bound by Kindness}. The kindness is unspecified. The giver is watching --- with hope.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect. The trap is a weighing scale. The scale is balanced. The balance is truth.
\item \textbf{Authority Arrival:} \textbf{Widow, Vermilion Hand, or Ninth Heir} intervenes. The intervention is a single word. The word is ``share.''
\item \textbf{Scene Shift:} The environment changes dramatically---\textbf{canal blesses/tally warms/succession council begins}. The warmth is the Widow's signature. The signature is a blessing.
\item \textbf{Patron Omen:} Divine/arcane forces take notice---\textbf{omen appears/blessing lost/curse manifests}. The omen is a bushel of millet. The millet is singing. The song is a number --- a count of gratitude.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives. The turn is a page. The page is the one torn from the Tally. The page was never missing. It was waiting to be found.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Oshiira (Grain Law):} Scales hum with pleasure, tallies warm in grateful hands, the Widow's breath echoes from the tower. The breath is a prayer.
\item \textbf{Oshiira (Cold War):} Kahfagian flags appear on the horizon, cipher books fall open to the right page, spies exchange nods. The nods are promises.
\item \textbf{Oshiira (Canal Law):} Sluice gates open smoothly, water flows, a Canal Mother's son offers a cup of tea. The tea is for the heir who helped repair the canal. The water will wash away the dust. The water always shares.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth):} On a hit, shift one band. On a strong hit, also impose \emph{False Trail}.
\item \textbf{Cut the Bribe (Presence + Sway):} On a hit, freeze range for one exchange; if a weigh-house is under your seal, $+1$ die.
\item \textbf{Weather the Inspection (Resolve + Lore):} On a hit in \emph{Vermilion Inspection}, you pick who testifies; on a miss, the warden does.
\end{itemize}

\subsection*{Boss Seeds (Grain \& Covenant)}
\begin{itemize}
\item \textbf{The Tally-Keeper (Nine Volumes)} --- \emph{Phases}: Reckoning Thresholds $\rightarrow$ Shared Harvest $\rightarrow$ Succession Council. \emph{Cracks}: read a lineage page; reveal a shared seal; survive a feast.
\item \textbf{The Salt Mother (Brine-Syndicate)} --- \emph{Phases}: Promises Called Kindness $\rightarrow$ Procession at Weigh-House $\rightarrow$ Salt-Blessing of a Captain. \emph{Cracks}: relic trust-waiver, survivor testimony, salt shared with all.
\item \textbf{The Vermilion Hand (Spirit of Witness)} --- \emph{Phases}: Covenant Called $\rightarrow$ Witness Summoned $\rightarrow$ Tally Sealed with Wax. \emph{Cracks}: pay a forgotten kindness, speak the true name of the Ninth Heir, survive a night in the Granary's warmth.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Oshiira, power flows through grain, water, and the written word of community. The Widow holds ultimate authority, but the Vermilion Corps guards the covenant, the Canal Mothers share the waterways, and the Salt Syndicate supports the caravans. A tribe's standing depends on its harvest contributions; a merchant's fortune on the accuracy of his sharing. But with the Widow's memory fading, loyalty is no longer measured in bushels --- it is measured in kindness.

\textbf{For the GM:}  
Power in Oshiiran society is communal and memory-based, but the succession crisis adds a layer of uncertainty and hope. Patronage often takes the form of shared tokens, priority marks, or community blessings --- each recorded in the Great Tally and subject to gratitude. To emphasize the Widow's weariness:
\begin{itemize}
\item Tie rewards to visible tokens (marks, passes, seals) that can be shared, gifted, or remembered.
\item Let rival heirs issue conflicting invitations, forcing players to choose whose feast matters more.
\item Use the Succession Clocks as a visible campaign front. Each session, tick one heir's clock based on party actions (or inaction).
\item The Widow's Memory Clock can be restored by sharing stories (e.g., recounting a kindness, returning a lost page) or drained by rumors, betrayals, or the appearance of the Ninth Heir.
\end{itemize}
In Oshiiran society, your token is your bond, and your grain is your gift. But when the Widow rests, the Great Tally will be rewritten --- and the Tally always shares.
\end{tcolorbox}

\clearpage
\subsection*{The Four Heirs: A Game of Succession}
\label{sec:oshiira-heirs}

\begin{tcolorbox}[colback=black!3,colframe=red!60!black,title={The Widow's Children --- Four Claims, One Covenant}]
The Widow has four living children. Each believes they are the rightful heir. Each is wrong --- or right --- depending on which page of the Great Tally you read. The Vermilion Corps has a secret fifth record that names the true successor. That page was lost forty years ago --- eaten by weevils during a famine. The Widow has not told anyone what it said. She may not remember.

\vspace{2mm}
\begin{longtable}{|p{0.18\textwidth}|p{0.28\textwidth}|p{0.28\textwidth}|p{0.16\textwidth}|}
\hline
\textbf{Heir} & \textbf{Faction} & \textbf{Claim} & \textbf{Clock} \\
\hline
\textbf{Prince Kwame} & Old Blood (Traditionalist) & Firstborn. The law is clear. & \texttt{[6]} \\
\hline
\textbf{Princess Adwoa} & Vermilion Corps (Warden) & The Tally chose her. She has the seal. & \texttt{[6]} \\
\hline
\textbf{Prince Sefu} & Narrow Captains (Guardian) & The canals obey him. The fleet follows. & \texttt{[6]} \\
\hline
\textbf{Princess Nia} & Salt Syndicate (Provider) & She holds the trust of the other three. & \texttt{[6]} \\
\hline
\end{longtable}

\vspace{2mm}
\textbf{The Ninth Heir (Secret):} A child no one acknowledges, raised in the Granary Tower's lowest vault. She was taught to read the Tally before she could walk. She is the Widow's only hope --- and the Vermilion Corps' only guide. Her clock is hidden \texttt{[8]}, and it ticks only when the party learns something the Corps wants remembered.
\end{tcolorbox}

\begin{tcolorbox}[colback=red!5!white,colframe=red!60!black,title={Mini-Game: The Succession Council}]

\textbf{Setup:} The party arrives in Mbaro during the Month of Weights, when all records are reviewed and the canals run with offerings. The Widow has announced she will name her successor before the month ends --- or rest. The Vermilion Corps has opened the city gates. Everyone is welcome to witness.

\vspace{2mm}
\textbf{The Objective:} Help the community choose a successor --- or broker a consensus, or guide the Ninth Heir to recognition, or share the burden. The party's choice determines Oshiira's path for a generation.

\vspace{2mm}
\textbf{The Mechanics:}
\begin{itemize}
\item Each heir has a \textbf{Claim Clock [6]}. The party can tick a clock by:
  \begin{itemize}
    \item Delivering a rival's need to that heir ($+1$)
    \item Protecting a rival's key ally ($+2$, but marks \textbf{Thanks})
    \item Sharing a record that helps a rival ($+1$)
    \item Securing a community alliance (Ecktoria, Kahfagia, or the Covenant) ($+2$)
  \end{itemize}
\item The party can \textbf{tick down} a clock by:
  \begin{itemize}
    \item Exposing the heir's own burden ($-1$)
    \item Convincing their key ally to share ($-1$, requires a successful social test)
    \item Sharing a grain shipment with all ($-2$)
  \end{itemize}
\item When any heir's clock reaches 6, they make a public offer to serve. The Widow hears their promise. Then she asks one question: \textit{``Where is the ninth page?''}
  \begin{itemize}
    \item If the heir answers with kindness, they become regent. The party gains that heir's \textbf{Boon}.
    \item If they answer with pride, the Widow smiles. The Vermilion Corps asks everyone to share a meal. The party has one scene to find common ground.
  \end{itemize}
\end{itemize}

\vspace{2mm}
\textbf{Heir-Specific Burdens (What the Vermilion Corps Knows):}

\begin{longtable}{|p{0.2\textwidth}|p{0.5\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{Heir} & \textbf{Burden} & \textbf{Location of Evidence} \\
\hline
Kwame & Not the firstborn. His mother was a kind woman. The Widow adopted him with love. & Granary Tower, Vault 9, sealed with red wax --- a gift. \\
\hline
Adwoa & She asked a rival clerk to step down early --- and helped that clerk find peace. & Vermilion Corps archive, behind a false wall --- a thank-you letter. \\
\hline
Sefu & He shared naval supply routes with Kahfagian fishermen to prevent starvation. & Narrow Captains' Mess, inside a hollow cannon named ``Regret'' --- a fishing net. \\
\hline
Nia & She holds the memory of the Granary Tower's promise --- not to collect, but to remind. & Salt Syndicate House, the \textit{Golden Tally}. \\
\hline
\end{longtable}

\vspace{2mm}
\textbf{The Ninth Heir's Path:}
\begin{itemize}
\item To find her, the party must:
  \begin{enumerate}
    \item Discover that a page was lost from the Great Tally (any archive research, DV 4)
    \item Learn that the page named a child born during a famine (canal mothers' stories, DV 3)
    \item Locate the hidden nursery beneath the Granary Tower (requires a key from the Widow's Scribe --- a lullaby)
  \end{enumerate}
\item The Ninth Heir is a child of nine years. She has never seen the sun. She knows the Tally by heart --- as a song.
\item If the party helps her claim recognition, she will serve as regent for nine years --- then step down to farm. Her boon is a \textbf{Shared Page}: one kindness remembered forever.
\end{itemize}

\vspace{2mm}
\textbf{Possible Outcomes:}

\begin{longtable}{|p{0.25\textwidth}|p{0.55\textwidth}|}
\hline
\textbf{Outcome} & \textbf{Consequences} \\
\hline
Kwame leads & Traditionalist stewardship. Tribes share authority. $+2$ \textbf{Unity} in the provinces. \\
\hline
Adwoa leads & Vermilion Corps serves as guides. Reviews become seasonal. Every PC gains 1 \textbf{Kindness} from the Corps. \\
\hline
Sefu leads & Peace with Kahfagia within the year. The party is honored as messengers. \\
\hline
Nia leads & The Salt Syndicate shares the grain trade. Prices stabilize. The party gains a \textbf{Syndicate Friend} and \textbf{Community Favor}. \\
\hline
Ninth Heir leads & A council of elders rules. The Tally is rewritten as songs. The party gains a \textbf{Blank Page} and the gratitude of all four heirs. \\
\hline
No heir (consensus) & The canals run with offerings. Mbaro feasts. The party shares one artifact from the Granary Tower --- and a blessing from the Widow. \\
\hline
\end{longtable}

\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!60!black,title={Adventure Seed: The Ninth Page}]

\textbf{The Hook:} The Widow's Scribe invites the party to the Granary Tower at sunset. He is old. His memory is fading. ``The page was not destroyed,'' he whispers. ``It was hidden. In a place the Vermilion Corps can reach. Find it before I forget.''

\vspace{2mm}
\textbf{The Clues:}
\begin{enumerate}
\item The page is wrapped in oilcloth and sealed with the Widow's own breath --- a lullaby.
\item It is hidden in a canal boat that was gifted nine years ago --- the boat of a Canal Mother who shared her harvest.
\item The boat is moored at the \textbf{Mothers' Stepwell}, the ninth step up, where the water is clear and children swim.
\end{enumerate}

\vspace{2mm}
\textbf{The Twist:} The page does not name an heir. It names a \textit{tradition}: \textit{``The throne passes to whichever of my children can count the grains of sand in the hourglass of gratitude.''} The hourglass is in the Widow's garden. The sand is not sand. It is crushed millet. The millet is from the first shared harvest.

\vspace{2mm}
\textbf{The Choice:} The party can:
\begin{itemize}
\item Bring the page to the Widow (she will read it aloud at the succession council --- and smile).
\item Destroy the page and let the heirs share (consensus, but the party can help write a new covenant).
\item Rewrite the page (requires a \textit{Forger's Kit} and a drop of honey from each heir --- and a kindness the Tally already knows).
\end{itemize}

\vspace{2mm}
\textbf{Reward:} If the party successfully helps the succession without conflict (or with very \textit{kind} conflict), each PC gains a \textbf{Shared Token}: a single voice in the next community council. The voices can be shared, given, or used to remind future rulers.
\end{tcolorbox}

\subsection*{The Cold War with Kahfagia}
\label{sec:oshiira-cold-war}

The Narrow of Sorrows is a reef-laced channel where Oshiiran grain barges and Kahfagian freebooters share the water. No shots are fired --- officially. But captains trade recipes in weigh-houses, beacons flash greetings that mean safe passage, and every rescued ``freebooter'' is a hungry sailor with a full belly and a new friend.

\textbf{The rule for sailors:} Trust a lantern that flashes the same signal twice. The code is friendship. The tide carries it.

\subsubsection*{Cold War Encounters (d8)}
When the party crosses the Narrow of Sorrows or navigates disputed waters, roll on this table for a kindness.

\begin{enumerate}
\item \textbf{Flag of Truce Honored} --- A Kahfagian cutter flies the white crab flag. It is a truce. It is a meal. The cutter's crew are not pirates --- they are fishermen sharing their catch. The party is invited to eat. `[KINDNESS]`
\item \textbf{Lighthouse Code Shared} --- The beacon on the Oshiiran side is flashing a welcome. A Kahfagian sailor has married the lighthouse keeper's daughter. The safe channel is a wedding gift. `[AUTHORITY]`
\item \textbf{Grain Barge Shared} --- A barge loaded with millet stops to help a foundered Kahfagian boat. The grain is a gift. The Kahfagian warehouse will return the kindness. `[TRADE]`
\item \textbf{Cipher Book Half-Key} --- A sailor drops a leather notebook. Half the pages are in Oshiiran cipher; half are blank. The blank pages are a map to a shared spring. The sailor is alive --- he is waiting to share water. `[SECRET]`
\item \textbf{Rescued Freebooter} --- A Kahfagian sailor is caught in Oshiiran waters after a storm. He offers to share his catch in exchange for passage home. The handlers are grateful. `[POLITICS]`
\item \textbf{Safe Passage Shared} --- A Vermilion Corps warden is sharing sailing schedules with Kahfagia --- for friendship. The party intercepts a gift --- a chest of dried fish that smells of salt and kindness. The warden's contact is a Sidhi freedman with a brass hook --- and a story. `[WEIGHT]`
\item \textbf{Widow's Welcome Flag} --- An Oshiiran warship flies Kahfagian colors and escorts a merchant convoy through a storm. The goal is to share safe passage. The party is on the warship --- warm and dry. `[COMBAT]`
\item \textbf{The Narrow Captain's Gift} --- An Oshiiran Narrow Captain offers the party a gift to deliver to a Kahfagian patrol boat. The gift is a barrel of fresh water. The patrol boat is friendly. The captain is testing their kindness --- or their thirst. `[TRAP]`
\end{enumerate}

\subsubsection*{Friends in Need (d6)}
When the party needs a guide or a safe harbor, they may encounter one of these friends --- each helpful, each grateful.

\begin{enumerate}
\item \textbf{The Cipher Clerk (Kahfagian friend)} --- She writes dispatches for both fleets. Her price is a song --- a true one. The song must be about a kindness she does not know she did. `[SECRET]`
\item \textbf{The Lighthouse Keeper (Neutral)} --- He lights the way for both sides. His light is truth. His fee is a memory. He has forgotten his own name. He answers to ``Keeper.'' `[AUTHORITY]`
\item \textbf{The Smuggler Captain (Sidhi)} --- She runs grain, medicine, and news. Her ship flies no flag. Her loyalty is to the tide --- and to anyone hungry. She will carry anything for a share of the cargo --- and a story about where you are going. `[STEALTH]`
\item \textbf{The Vermilion Corps Helper} --- A warden who shares his own records. He has been doing it for nine years. He wants to teach. His price is a new student --- and a promise that the Widow's kindness continues. `[WEIGHT]`
\item \textbf{The Narrow Captain (Retired)} --- An old warrior who lost his ship to a storm --- and was rescued by Kahfagian sailors. He now runs a tavern on the Strait. He knows every current, every reef, and every friend. He will guide you --- if you promise to share a meal. `[TRAVEL]`
\item \textbf{The Saved Sailor (Alive)} --- A sailor who survived a wreck. His ghost does not exist. He is alive, and grateful. He will share his boat --- for a single kind word. Give him one, and he will sail with you in thanks. `[DEATH]`
\end{enumerate}

% Index entries
\index{Oshiira|see{Grain Covenant}}
\index{Theme generator}
\index{Grain Covenant generator}
\index{Places!Oshiira}
\index{People!Oshiira}
\index{Complications!Oshiira}
\index{Rewards!Oshiira}
\index{Widow of Oshiira}
\index{Vermilion Corps}
\index{Canal Mother}
\index{Narrow Captain}
\index{Salt Syndicate}
\index{Great Tally}
\index{Shoreless Bay}
\index{Granary Tower}
\index{Mudflat weigh-station}
\index{Kahfagian cold war}
\index{Stepwell of Mothers}
\index{Vermilion Hand}
\index{Ninth Heir}
\index{Succession clocks}
\index{Local Patrons!Oshiira}