\section{Sekogo ‚àí‚àí‚àí ``The Green Council''}
≈Çabel{chap:sekogo}

\textbf{Key Demographics: } Humans
\textbf{Capital City: } Ukwe (The City)

\section*{Region Diagnostic Template}

\subsection*{\textbf{\textsf{Sekogo}}}

\paragraph{Current Hook (one sentence)}
\framebox[≈Çinewidth][l]{%
\begin{minipage}{≈Çinewidth}
\vspace{2mm}
The Ukwe (Speaking Tree) demands a name: a caravan master broke his oath to the grove, and the roots are hunting him down.
\vspace{2mm}
\end{minipage}}

\paragraph{Core Conflict (one sentence)}
\framebox[≈Çinewidth][l]{%
\begin{minipage}{≈Çinewidth}
\vspace{2mm}
Players are caught between the Conservators (who demand stasis) and the Wakers (who demand growth through fever and blood), while the Leopard Society hunts anyone who breaks the balance.
\vspace{2mm}
\end{minipage}}

\paragraph{The One Thing Only This Region Does}
\framebox[≈Çinewidth][l]{%
\begin{minipage}{≈Çinewidth}
\vspace{2mm}
Promises are physical objects (\textit{mukanda} bark‚àícords) that tighten when strained and rot when released; breaking them attracts the \textit{Mwitu} (Hollow).
\vspace{2mm}
\end{minipage}}

\paragraph{Interactive Mechanics (player‚àífacing)}
\begin{itemize}
‚Ñπtem[\card{‚ôÝ}{1}] \textbf{The Ukwe's Witness:} Speak a secret to a marked tree to gain \textbf{1 Boon}, but the tree remembers it forever (GM gains \textbf{1 SB} when that secret is relevant).
‚Ñπtem[\card{‚ô•}{2}] \textbf{Honey‚àíTithe:} Offer honey to gain safe passage; refusing marks \textbf{1 Obligation} to the Grove Spirit.
‚Ñπtem[\card{‚ô£}{3}] \textbf{Mask Law:} Wearing a ceremonial mask grants authority but binds you to its taboos; breaking them causes the mask to crack (\textbf{‚àí‚àí1 die} social).
‚Ñπtem[\card{‚ô¶}{4}] \textbf{Ananse's Bargain:} Renegotiate a promise once per session (Spirit + Lore, DV 4); success shifts terms, failure doubles the burden.
\end{itemize}

\paragraph{Campaign Timers (GM‚àífacing)}
\begin{itemize}
‚Ñπtem \textbf{Waker Fever [6]} ‚Äì \textit{Advances when:} Players ignore spirit warnings or use Waker spores. \textit{At full:} The jungle becomes hostile; navigation rolls at \textbf{‚àí‚àí1 die}.
‚Ñπtem \textbf{Ukwe's Memory [8]} ‚Äì \textit{Advances when:} Players break promises or cut marked trees. \textit{At full:} The \textit{Mwitu} manifests; the party is lost until a debt is paid.
\end{itemize}

\paragraph{Faction Triad}
\begin{tabular}{|p{0.2≈Çinewidth}|p{0.25≈Çinewidth}|p{0.25≈Çinewidth}|p{0.2≈Çinewidth}|}
‚Ñéline
\textbf{Faction} & \textbf{Goal} & \textbf{Method} & \textbf{Why players care} \\
‚Ñéline
Leopard Society & Restore Balance & Hunt oath‚àíbreakers and slavers & They offer protection but demand justice. \\
‚Ñéline
Earth Society & Preserve Memory & Keep the Calendar of Witness & They hold the records of past debts. \\
‚Ñéline
Wakers & Awaken the Jungle & Use fever‚àímists and blood rites & They offer power but risk corruption. \\
‚Ñéline
\end{tabular}

\paragraph{The Price}
\framebox[≈Çinewidth][l]{%
\begin{minipage}{≈Çinewidth}
\vspace{2mm}
Power requires memory loss, Obligation to spirits, or physical marks (scars, thorns); honey is currency, but the bees remember who took it.
\vspace{2mm}
\end{minipage}}

\paragraph{Geography as Constraint}
\framebox[≈Çinewidth][l]{%
\begin{minipage}{≈Çinewidth}
\vspace{2mm}
The jungle is a witness; marked trees cannot be cut, and paths shift if promises are broken. The Crimson Valley bridges only hold if web‚àílaw is honored.
\vspace{2mm}
\end{minipage}}

\paragraph{The Ninth Taboo manifestation}
\framebox[≈Çinewidth][l]{%
\begin{minipage}{≈Çinewidth}
\vspace{2mm}
The Ninth Root searches for erased names; the ninth knot on a \textit{mukanda} is left loose for the Hollow; the ninth bell of Ananse's Child rings when a promise is broken.
\vspace{2mm}
\end{minipage}}

\paragraph{Current Weaknesses (honest self‚àíassessment)}
\begin{itemize}
‚Ñπtem Too many spirit types (\textit{Mzuka}, \textit{Kizimba}, \textit{Ngozi}, etc.) may confuse players.
‚Ñπtem Faction goals overlap too much (all care about promises).
‚Ñπtem Risk of "magic everywhere" diluting the cost of spellcasting.
\end{itemize}

\vspace{1cm}
‚Ñérule
\vspace{0.5cm}
\noindent\emph{Design Fixes Applied: Consolidated spirits into three categories (Ancestral, Nature, Vengeant). Clarified faction distinctions (Stasis vs. Growth vs. Enforcement). Integrated core \textbf{Fate's Edge} mechanics (Boons, SB, Obligation) into all rewards.}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The jungle remembers every promise. The mask is the law. The honey is a bond.}

\begin{quote}
  \textbf{Leopard Society Hunter (Elite):}  
  ``The jungle has a memory. When a fever rises, we cut it out. When an oath is broken, we restore the balance. We are not cruel. We are necessary. The tree remembers every deed, and we are its teeth.''
\end{quote}

\begin{quote}
  \textbf{Honey‚àíGatherer (Commoner):}  
  ``The bees know my name. They do not sting me because I sing to them the song of the first blossom. My grandmother taught me that the honey is not a gift; it is a trust. One day, the grove will need something from you, and you will remember the taste.''
\end{quote}

\textbf{Genre / Mood:} Jungle folk horror, spirit diplomacy, mask‚àílaw.

\textbf{Starting Location:} A honey‚àíorchard at the jungle's edge, where the trees hum and the air is thick with bees. A masked figure in a leopard pelt steps from the shadows: ``The Ukwe has asked for a name. A caravan master broke his oath to the grove. Find him before the roots do ‚àí‚àí‚àí or become his witness.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme & Atmosphere}]
Sekogo lies between the Sahel grasslands and the southern plateau ‚àí‚àí‚àí a dry savannah that grows wetter as you travel east, until the baobabs give way to deep jungle and the air turns thick with the smell of damp earth and overripe fruit. Here, the Green Council rules through consensus, honey‚àíorchards are sacred, and the old spirits still walk. Slavery is forbidden; the jungle itself would rise against any slaver who crossed its border. But the Council is divided: Conservators want to protect the old pacts, while Wakers argue that the jungle has grown sluggish and needs to be roused with fever‚àímists, thorn‚àíspirits, and older, bloodier rites. At the eastern edge lies the Crimson Valley, home to the Lethai‚àíar ‚àí‚àí‚àí chestnut‚àíbrown elves whose silk bridges span red canyons and whose web‚àílaw keeps both empires from using the valley as a shortcut to war.

\vspace{2mm}
\textbf{Neighbors and the Unclaimed Ground:}
\begin{itemize}
  ‚Ñπtem \textit{North ‚àí‚àí Ameria and the Neutral Kingdoms:} Beyond Sekogo's northern savannah lies a patchwork of small, independent kingdoms ‚àí‚àí‚àí buffer states that pay tribute to neither Sekogo nor Kahfagia. Their courts are weak, their laws a creole of jungle custom and coastal tariff. Smugglers and exiles thrive there. The Leopard Society occasionally hunts across the border, but always with a local guide.
  ‚Ñπtem \textit{East ‚àí‚àí The Crimson Valley (Lethai‚àíar):} The Ukwe's roots extend into the valley's red stone. The Lethai‚àíar silk‚àíwardens respect the Speaking Tree, and the Green Council's witnessed promises are sometimes recorded on a Lethai‚àíar cord. A bridge toll paid in Sekogo honey is honored on both sides of the gorge.
  ‚Ñπtem \textit{South and West ‚àí‚àí The Jungle's Own:} Sekogo's borders are not lines but gradients. The deeper you go, the more the Ukwe's roots thicken and the Leopard Society's law becomes the only law.
\end{itemize}

\vspace{2mm}
\textbf{What you notice first:} The sound of bees ‚àí‚àí‚àí a constant, low drone that never stops, even at night. The way the trees lean toward you when you speak a secret. The smell of honey and rotting fruit, sweet and sharp at once.

\textbf{The one rule that kills newcomers:} Never cut a marked tree. Each such tree is a living witness. The bark grows over the names of those who have broken harmony, but the wood remembers. Cut it, and the Ukwe will add your name to its memory.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Root}

Power here is measured in witnessed promises, honey tithes, and the balance between jungle and council. The Leopard Society hunts those who break the deepest taboos ‚àí‚àí‚àí oath‚àíbreakers, slavers, outsiders who take more than they give. The Earth Society keeps the \textbf{Calendar of Witness}, a sacred list written on bark and buried at the Ukwe's roots, preserving memories that must not be forgotten.

The spirits of Sekogo are consolidated into three kinds for play: \textit{Ancestral} (bound to groves, speak in dead voices), \textit{Nature} (guardians like crocodiles or baobabs, show rather than speak), and \textit{Vengeant} (born from blood‚àípromises, hunt until answered). The Hollow manifests here as the \textit{Mwitu}, the forest that forgets where the village ends ‚àí‚àí‚àí and that forgetfulness is a hunger.

\subsection*{Prominent NPCs of Sekogo}
≈Çabel{sec:sekogo‚àínpcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
‚Ñéline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
‚Ñéline
Hunter Okoro & Leopard Society mask‚àíwearer & Hunt down a Waker who poisoned a foreign caravan & He never removes his mask. The mask is his real face now. \\
‚Ñéline
Elder Imani & Earth Society Ogboni & Keep the Calendar of Witness from being forgotten & She speaks only while holding a piece of the Ukwe's bark. She cannot lie. \\
‚Ñéline
Nganga Malika & Spirit‚àídoctor of the \textit{minkisi} & Heal a wound caused by a broken \textit{nkisi} & She carries a carved figure with nine nails. The ninth nail is loose. \\
‚Ñéline
Waker Sefu & Radical who wants to wake the jungle & Poison the wells of a Conservator village & A green shoot grows from his shoulder. It bleeds when he lies. \\
‚Ñéline
Lethai‚àíar Silk‚àíWarden Anara & Guard of the Crimson Valley bridges & Keep the web‚àílaw intact against Oshiiran pressure & Her skin‚àítattoos record a witnessed promise her grandmother carried. She will not say to whom. \\
‚Ñéline
The Ukwe's Voice & A spirit that speaks through a cracked root & Remind the Council of a promise nine years overdue & It speaks in the voice of a dead child. The child was a witness. \\
‚Ñéline
Ananse's Child (Trickster) & A spider‚àíspirit in human form & Unravel oaths for the joy of chaos & It wears nine tiny bells on its ankles; each bell is a broken promise. \\
‚Ñéline
Mami Wata & Water spirit of the Crimson Falls & Collect a memory from every soul that crosses her pool & Her hair is black water; her eyes are river stones. She never blinks. \\
‚Ñéline
\end{longtable}

\subsection*{Names of Sekogo}
≈Çabel{sec:sekogo‚àínames}

Inspired by the savannah and jungle regions: melodic, with nasal consonants and tonal shifts. Surnames often denote a totem animal, a grove, or a witnessed deed.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
‚Ñéline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
‚Ñéline
Okoro & Imani & Ngoma & \textit{the Masked} \\
‚Ñéline
Sefu & Malika & Ukwe‚àíborn & \textit{the Honey‚àíTongue} \\
‚Ñéline
Chima & Anara & Leopard‚àíkin & \textit{the Oath‚àíKeeper} \\
‚Ñéline
Kofi & Zola & Root‚àíwalker & \textit{the Ninth Nail} \\
‚Ñéline
Tunde & Amara & Grove‚àísinger & \textit{the Unmarked} \\
‚Ñéline
Biko & Nkiru & Minkisi‚àíhand & \textit{the Still Witness} \\
‚Ñéline
Adofo & Adaeze & Ananse‚àífo & \textit{the Trickster's Child} \\
‚Ñéline
\end{longtable}

\paragraph*{(Orchard/Circle/Bridge)} Honey‚àíorchard where the trees hum; druid circle with moss‚àícovered stones and a leopard skull at center; Lethai‚àíar silk bridge spanning a crimson gorge with the Ukwe's roots clutching the water below.

\section*{Spades ‚àí‚àí‚àí Places}
≈Çabel{sec:sekogo‚àíplaces}
\begin{enumerate}
\setcounter{enumi}{1}
‚Ñπtem \textbf{Honey‚àíOrchard} ‚àí‚àí‚àí The trees hum. The air is thick with bees.
  ‚Ñéfill \textit{The honey is dark red. It tastes of iron and forgotten promises. The ninth hive is always empty ‚àí‚àí‚àí the bees say the Hollow has taken the queen.} \texttt{[SPIRIT]}
‚Ñπtem \textbf{Druid Circle (Leopard Skull)} ‚àí‚àí‚àí A ring of moss‚àícovered stones. A skull sits at the center.
  ‚Ñéfill \textit{The skull is warm. When the Leopard Society hunts, the eye sockets glow ‚àí‚àí‚àí and the \textit{mzuka} of the skull whispers the quarry's true name.} \texttt{[RITUAL]}
‚Ñπtem \textbf{Lethai‚àíar Silk Bridge} ‚àí‚àí‚àí A span of woven silk across a crimson gorge.
  ‚Ñéfill \textit{The silk is made from spider‚àísap and oath‚àíhair. It will not break unless an oath is called. The ninth strand is invisible ‚àí‚àí‚àí it is the Hollow's witness.} \texttt{[OATH]}
‚Ñπtem \textbf{The Ukwe (Speaking Tree)} ‚àí‚àí‚àí A baobab so old its bark is carved with a thousand names.
  ‚Ñéfill \textit{The tree speaks in the voices of the dead. It never gives answers ‚àí‚àí‚àí only new questions and old memories. The ninth root never stops growing; it searches for the name that was erased.} \texttt{[WITNESS]}
‚Ñπtem \textbf{Nkisi Grove} ‚àí‚àí‚àí A clearing where broken power‚àíobjects are buried.
  ‚Ñéfill \textit{The ground hums. The nails that sealed the \textit{minkisi} have grown into iron thorns. The ninth thorn is always warm; it marks the grave of the nganga who drove the first nail.} \texttt{[MAGIC]}
‚Ñπtem \textbf{Calendar of Witness (Earth Society Vault)} ‚àí‚àí‚àí A hollow log sealed with wax and bone.
  ‚Ñéfill \textit{Inside are strips of bark. Each strip records a broken harmony. The bark is still growing ‚àí‚àí‚àí and the ninth strip has no writing. It waits for a promise that has not yet been broken.} \texttt{[MEMORY]}
‚Ñπtem \textbf{Waker's Fever‚àíMist Pool} ‚àí‚àí‚àí A stagnant pond that steams even in winter.
  ‚Ñéfill \textit{The mist smells of rotting flowers. Breathe it, and you will see the jungle as the Wakers see it ‚àí‚àí‚àí a place that is starving, waiting to be woken. The ninth breath of the mist shows the \textit{Mwitu}.} \texttt{[WEATHER]}
‚Ñπtem \textbf{Leopard Society Lodge} ‚àí‚àí‚àí A longhouse hidden in a thicket of thorn‚àívines.
  ‚Ñéfill \textit{The walls are made of human jawbones. The jaws are from oath‚àíbreakers. They still click when a new name is spoken ‚àí‚àí‚àí and the ninth click is always the name of the next quarry.} \texttt{[SECRET]}
‚Ñπtem \textbf{Crimson Valley Gorge} ‚àí‚àí‚àí A canyon of red stone, crossed by silk bridges at three points.
  ‚Ñéfill \textit{The stone is warm. It bleeds rust when the web‚àílaw is broken. The ninth bridge is a myth ‚àí‚àí‚àí it appears only when a promise is witnessed by a ghost.} \texttt{[THRESHOLD]}
‚Ñπtem[J] \textbf{Conservator Archive} ‚àí‚àí‚àí A cave filled with sealed clay pots. Each pot contains a memory.
  ‚Ñéfill \textit{The memories are for sharing. The price is a memory of equal weight. The ninth pot is cracked ‚àí‚àí‚àí the memory inside is leaking, and no one remembers what it was.} \texttt{[MEMORY]}
‚Ñπtem[Q] \textbf{The Grove of Sleeping Spirits} ‚àí‚àí‚àí A forest where the \textit{minkisi} go to die.
  ‚Ñéfill \textit{The spirits do not wake. They dream of the first nail. The dream is contagious ‚àí‚àí‚àí and the ninth dream is the Hollow's own.} \texttt{[DEATH]}
‚Ñπtem[K] \textbf{The Ukwe's Root‚àíChamber} ‚àí‚àí‚àí A cavern beneath the tree where the roots form a throne.
  ‚Ñéfill \textit{The throne is made of braided roots. Sit on it, and the tree will ask one question. Answer poorly, and you never leave. Answer with a lie, and the ninth root will bind your tongue.} \texttt{[JUDGMENT]}
‚Ñπtem[A] \textbf{The Crimson Falls} ‚àí‚àí‚àí A waterfall that runs red after the first rain of spring.
  ‚Ñéfill \textit{The water is not blood. It is rust from the canyon walls. The Lethai‚àíar say it is the earth's memory of a war ‚àí‚àí‚àí and the ninth drop is always cold, because it fell on a broken oath.} \texttt{[SPIRIT]}
‚Ñπtem[A+] \textbf{The Pool of Mami Wata} ‚àí‚àí‚àí A still, black pool behind the Crimson Falls, reachable only by diving underwater.
  ‚Ñéfill \textit{The water is warm. It shows not your reflection, but the face of the person whose memory the spirit holds. The ninth ripple is always your own face ‚àí‚àí‚àí but older, and weeping.} \texttt{[WATER]}
\end{enumerate}

\paragraph*{(Gatherer/Druid/Warden)} Honey‚àígatherer with a fruit offering and a leopard claw necklace; Green Council druid with a thorn‚àístaff and milky eyes that see the weight of promises; Lethai‚àíar silk‚àíwarden, masked and quiet, his mask painted with the Ukwe's roots.

\section*{Hearts ‚àí‚àí‚àí People & Factions}
≈Çabel{sec:sekogo‚àípeople}
\begin{enumerate}
\setcounter{enumi}{1}
‚Ñπtem \textbf{Honey‚àíGatherer} ‚àí‚àí‚àí She wears a leopard claw around her neck. The claw is from a cat she killed with her bare hands.
  ‚Ñéfill \textit{She offers a fruit to the trees before each harvest. The fruit is always a little rotten ‚àí‚àí‚àí the ninth bite is for the \textit{Mwitu}.} \texttt{[COMMUNITY]}
‚Ñπtem \textbf{Green Council Druid (Conservator)} ‚àí‚àí‚àí Milk‚àíwhite eyes, a staff of living wood.
  ‚Ñéfill \textit{She cannot see the living. She sees only the threads of promises ‚àí‚àí‚àí glowing connections between every person and the Ukwe. The ninth thread is always frayed.} \texttt{[JUSTICE]}
‚Ñπtem \textbf{Waker Radical} ‚àí‚àí‚àí Carries a bag of fever‚àímist spores. His eyes are yellow.
  ‚Ñéfill \textit{He plants thorns in the paths of Conservator caravans. The thorns grow into walls overnight ‚àí‚àí‚àí and the ninth thorn is always a \textit{nkisi} nail, driven into a forgotten promise.} \texttt{[DANGER]}
‚Ñπtem \textbf{Lethai‚àíar Silk‚àíWarden} ‚àí‚àí‚àí Mask carved from the Ukwe's root, painted with oath‚àínames.
  ‚Ñéfill \textit{He never speaks above a whisper. The bridge hears him anyway ‚àí‚àí‚àí and the ninth whisper is always the name of someone who will break a promise before dawn.} \texttt{[OATH]}
‚Ñπtem \textbf{Leopard Society Hunter} ‚àí‚àí‚àí Wears a pelt that shifts and moves as if alive.
  ‚Ñéfill \textit{He can transform into a leopard at night. Or so he claims. No one has survived a hunt to confirm it ‚àí‚àí‚àí but the ninth transformation is said to be permanent.} \texttt{[FEAR]}
‚Ñπtem \textbf{Earth Society Elder} ‚àí‚àí‚àí Holds a piece of the Ukwe's bark. The bark is warm.
  ‚Ñéfill \textit{She asks every visitor: ``What have you forgotten to honor?'' The answer is carved into her memory ‚àí‚àí‚àí and the ninth answer is always the same: a name that cannot be spoken.} \texttt{[TRUTH]}
‚Ñπtem \textbf{Nganga (Spirit‚àíDoctor)} ‚àí‚àí‚àí Carries a \textit{nkisi} carved with nine nails. The ninth nail is loose.
  ‚Ñéfill \textit{She hammers a nail into the figure to call a spirit. Each nail is a cost. She has nine costs left ‚àí‚àí‚àí and the ninth nail, if driven, will call the Hollow.} \texttt{[MAGIC]}
‚Ñπtem \textbf{Oath‚àíKeeper (Hollow Manifestation)} ‚àí‚àí‚àí A figure made of bark and silence. It appears when the Calendar of Witness grows heavy.
  ‚Ñéfill \textit{It does not speak. It points at the one who broke harmony. That person's name appears on the nearest tree ‚àí‚àí‚àí and the ninth letter of the name glows green.} \texttt{[DEATH]}
‚Ñπtem \textbf{Lethai‚àíar Mark‚àíCutter} ‚àí‚àí‚àí A tattooist who records promises on skin.
  ‚Ñéfill \textit{She uses a needle of silk‚àífiber and ink made from crushed crimson stone. The ink glows when a promise is broken ‚àí‚àí‚àí and the ninth mark she ever cut was her own.} \texttt{[MEMORY]}
‚Ñπtem[J] \textbf{The Ukwe's Voice (Spirit)} ‚àí‚àí‚àí A crack in the tree's bark that speaks.
  ‚Ñéfill \textit{The voice changes. Sometimes it is a child. Sometimes it is a jarl. Sometimes it is the listener's own. The ninth word it speaks is always the name of a promise you forgot to make.} \texttt{[SPIRIT]}
‚Ñπtem[Q] \textbf{The First Nail (Legend)} ‚àí‚àí‚àí A mythic nganga who sealed a spirit into the Ukwe itself.
  ‚Ñéfill \textit{She is said to still walk the jungle, carrying a hammer made of root and bone. The ninth nail she drove was never found ‚àí‚àí‚àí it is still searching for the promise it was meant to seal.} \texttt{[LEGEND]}
‚Ñπtem[K] \textbf{The Leopard King} ‚àí‚àí‚àí A spirit that leads the Society's midnight hunts.
  ‚Ñéfill \textit{It has no mask. Its face is a leopard's. Its eyes are human. The ninth hunter who follows it never returns ‚àí‚àí‚àí they become the Leopard King's shadow.} \texttt{[FEAR]}
‚Ñπtem[A] \textbf{The Crimson Valley's First Weaver} ‚àí‚àí‚àí A ghost who teaches the web‚àílaw to new wardens.
  ‚Ñéfill \textit{She appears at dawn on the central bridge. She asks one question: ``Who watches the watchers?'' The ninth warden who answers wrongly is woven into the bridge.} \texttt{[JUSTICE]}
‚Ñπtem[A+] \textbf{Ananse's Child (Trickster)} ‚àí‚àí‚àí A small figure with too‚àílong fingers, wearing nine tiny bells.
  ‚Ñéfill \textit{It offers to ``renegotiate'' any promise ‚àí‚àí‚àí but its renegotiation always adds a new, hidden clause. The clause is a joke. The joke is always on you ‚àí‚àí‚àí and the ninth bell rings only when the joke is already done.} \texttt{[TRICKERY]}
\end{enumerate}

\paragraph*{(Spirit/Offer/Name)} A thorn‚àíspirit blocks the path ‚àí‚àí‚àí must offer honey or fight; the Ukwe demands a name before letting you pass; a Lethai‚àíar silk‚àíwalker challenges you to recite the web‚àílaw.

\section*{Clubs ‚àí‚àí‚àí Complications/Threats}
≈Çabel{sec:sekogo‚àícomplications}
\begin{enumerate}
\setcounter{enumi}{1}
‚Ñπtem \textbf{Thorn‚àíSpirit Block} ‚àí‚àí‚àí A wall of living thorns grows across the path.
  ‚Ñéfill \textit{The spirit ‚àí‚àí‚àí a \textit{kizimba} of the hedge ‚àí‚àí‚àí demands a name. Give a false one, and the thorns grow inward. Give a true one, and the thorns will part ‚àí‚àí‚àí but the spirit will remember your face.} \texttt{[SPIRIT]} \texttt{[OBLIGATION]}
‚Ñπtem \textbf{Ukwe Demands a Name} ‚àí‚àí‚àí The tree speaks: ``I have forgotten a promise. Remind me.''
  ‚Ñéfill \textit{If you cannot name a broken promise, the tree will take your name instead ‚àí‚àí‚àí and the ninth root will grow a new leaf with your face on it.} \texttt{[WITNESS]} \texttt{[TRUTH]}
‚Ñπtem \textbf{Lethai‚àíar Web‚àíLaw Challenge} ‚àí‚àí‚àí A silk‚àíwarden blocks the bridge: ``Recite the ninth clause.''
  ‚Ñéfill \textit{The ninth clause is unwritten. It changes with each warden. You must guess ‚àí‚àí‚àí but the ninth guess is always wrong, because the clause is the warden's own secret.} \texttt{[OATH]}
‚Ñπtem \textbf{Leopard Society Hunt} ‚àí‚àí‚àí A masked hunter appears at dusk. ``You are called to witness.''
  ‚Ñéfill \textit{Refusal earns a Leopard's Mark. The jungle becomes hostile to you ‚àí‚àí‚àí and the ninth beast you meet will be wearing a mask of your face.} \texttt{[FEAR]}
‚Ñπtem \textbf{Nkisi Hunger} ‚àí‚àí‚àí A spirit‚àíobject you carry begins whispering. It wants a memory.
  ‚Ñéfill \textit{Refuse, and the \textit{nkisi} shatters. The spirit inside is not friendly ‚àí‚àí‚àí and the ninth shard will embed itself in your shadow, following you until you feed it.} \texttt{[MAGIC]}
‚Ñπtem \textbf{Waker Poisoning} ‚àí‚àí‚àí A well is found with a green film on the water.
  ‚Ñéfill \textit{The poison is not lethal. It makes you dream of the jungle's hunger ‚àí‚àí‚àí and the ninth dream is always the \textit{Mwitu} asking, ``Why are you still here?''} \texttt{[WEATHER]}
‚Ñπtem \textbf{Earth Society Reckoning} ‚àí‚àí‚àí An elder demands to see what you carry in your memory.
  ‚Ñéfill \textit{The record is written in scars on your skin. You have no scars. That is suspicious ‚àí‚àí‚àí so the elder will give you nine scars, one for each forgotten promise you cannot name.} \texttt{[TRUTH]}
‚Ñπtem \textbf{Honey‚àíTithe Demanded} ‚àí‚àí‚àí A grove spirit appears: ``Pay the ninth jar.''
  ‚Ñéfill \textit{You have no honey. The spirit offers a bargain. The price is a story ‚àí‚àí‚àí and the ninth story you tell will become a new root in the Ukwe's memory.} \texttt{[TRADE]} \texttt{[OBLIGATION]}
‚Ñπtem \textbf{Crimson Valley Border Dispute} ‚àí‚àí‚àí Oshiiran surveyors cross the web‚àíline. The Lethai‚àíar arm their bows.
  ‚Ñéfill \textit{The party is caught between. Whoever you side with, the other will remember ‚àí‚àí‚àí and the ninth bridge will close to you forever.} \texttt{[POLITICS]}
‚Ñπtem[J] \textbf{The Ukwe's Root Breaches} ‚àí‚àí‚àí A root bursts through a village floor. It is looking for a specific name.
  ‚Ñéfill \textit{The name is carved on the root. The root is weeping. The one who broke the promise is still alive ‚àí‚àí‚àí and the ninth night they sleep, the root will find them.} \texttt{[DEATH]}
‚Ñπtem[Q] \textbf{Leopard Society Coup} ‚àí‚àí‚àí The Wakers have infiltrated the Lodge. The masks are fighting.
  ‚Ñéfill \textit{You cannot tell who is who. Both sides wear leopard pelts. Both sides smell of blood ‚àí‚àí‚àí and the ninth mask you see will be the one that speaks with the Leopard King's voice.} \texttt{[CONFLICT]}
‚Ñπtem[K] \textbf{Calendar Fire} ‚àí‚àí‚àí A spark catches the Earth Society vault. The memories are burning.
  ‚Ñéfill \textit{The smoke forms faces. The faces are those who broke harmony. They are screaming ‚àí‚àí‚àí and the ninth scream is your own, from a promise you made and then forgot.} \texttt{[DISASTER]}
‚Ñπtem[A] \textbf{The Crimson Falls Run Red} ‚àí‚àí‚àí The waterfall turns to blood. The Lethai‚àíar say the web‚àílaw has been broken.
  ‚Ñéfill \textit{The break was centuries ago. The falls are just catching up ‚àí‚àí‚àí and the ninth drop of blood carries the name of the first oath‚àíbreaker.} \texttt{[OMEN]}
‚Ñπtem[A+] \textbf{Mami Wata's Gaze} ‚àí‚àí‚àí A figure rises from the black pool behind the falls. It asks: ``What promise have you never been able to keep?''
  ‚Ñéfill \textit{If you lie, the water will take your voice. If you tell the truth, it will ask for a memory as price ‚àí‚àí‚àí and the ninth memory you offer will be the one you need most.} \texttt{[SPIRIT]} \texttt{[OBLIGATION]}
\end{enumerate}

\paragraph*{(Favor/Trust/Shard)} Leopard's Favor (once, call on the Leopard Society for a hunt ‚àí‚àí‚àí you are bound to them for a future hunt); Ukwe Memory (reroll a failed Survival test once per session, but GM gains 1 SB for a jungle complication); Nkisi Shard (add $+1$ Effect to a magical working, but the shard shatters and costs 1 Corruption).

\section*{Diamonds ‚àí‚àí‚àí Rewards/Leverage}
≈Çabel{sec:sekogo‚àírewards}
\begin{enumerate}
\setcounter{enumi}{1}
‚Ñπtem \textbf{Leopard's Favor} ‚àí‚àí‚àí A leopard claw wrapped in red thread. One use to summon a Society hunt.
  ‚Ñéfill \textit{The favor is a bond. The Society will remember. The price is a future hunt ‚àí‚àí‚àí and the ninth hunt will be for your own name, if you break the bond.} \texttt{[BOND]}
‚Ñπtem \textbf{Ukwe Memory (Leaf)} ‚àí‚àí‚àí A single leaf from the Speaking Tree. Reroll a failed Survival test once per session.
  ‚Ñéfill \textit{The leaf is green. It will not wilt. The tree remembers who took it ‚àí‚àí‚àí and the ninth time you use it, the leaf will turn black and whisper a name you have forgotten.} \texttt{[MEMORY]}
‚Ñπtem \textbf{Nkisi Shard} ‚àí‚àí‚àí A piece of a broken power‚àíobject. Add $+1$ Effect to a magical working, then it shatters.
  ‚Ñéfill \textit{The shard is warm. It whispers secrets in a language you almost understand ‚àí‚àí‚àí and the ninth secret is always about a promise you made to yourself and broke.} \texttt{[MAGIC]}
‚Ñπtem \textbf{Honey‚àíJar (Royal Tithe)} ‚àí‚àí‚àí A clay pot of dark red honey. Trade it for safe passage through any grove.
  ‚Ñéfill \textit{The honey is thick. It moves in the jar as if alive. Do not open it ‚àí‚àí‚àí the ninth time you open it, the bees will remember your name and ask for a tithe.} \texttt{[TRADE]}
‚Ñπtem \textbf{Earth Society Tablet} ‚àí‚àí‚àí A strip of bark with a broken promise recorded. Present it to recall a forgotten harmony.
  ‚Ñéfill \textit{The bark is still growing. The memory is still fresh ‚àí‚àí‚àí and the ninth time you use it, the bark will grow over your hand and show you the promise you yourself broke.} \texttt{[MEMORY]}
‚Ñπtem \textbf{Lethai‚àíar Silk Cord} ‚àí‚àí‚àí A braided cord that records a witnessed promise. Use it to enforce any oath.
  ‚Ñéfill \textit{The cord tightens when the oath is broken. It will not stop until blood is drawn ‚àí‚àí‚àí and the ninth drop of blood will be your own, if you witnessed a false promise.} \texttt{[OATH]}
‚Ñπtem \textbf{Leopard Mask (Temporary)} ‚àí‚àí‚àí Wear it to pass as a Society hunter for one scene.
  ‚Ñéfill \textit{The mask has eye holes that see the weight of promises. It also sees your own ‚àí‚àí‚àí and the ninth time you wear it, the mask will not come off.} \texttt{[SECRET]}
‚Ñπtem \textbf{Waker's Spore Pouch} ‚àí‚àí‚àí A bag of fever‚àímist spores. Release them to confuse pursuers.
  ‚Ñéfill \textit{The spores cause vivid hallucinations. The hallucinations are always of the Ukwe ‚àí‚àí‚àí and the ninth hallucination will be a question: ``Why did you lie?''} \texttt{[STEALTH]}
‚Ñπtem \textbf{Nganga's Hammer} ‚àí‚àí‚àí A small iron hammer. Drive a nail into any object to seal a spirit inside.
  ‚Ñéfill \textit{The hammer has a cost. Each nail requires a memory ‚àí‚àí‚àí and the ninth nail you drive will seal a fragment of your own soul into the object.} \texttt{[MAGIC]}
‚Ñπtem[J] \textbf{The Ukwe's Answer} ‚àí‚àí‚àí A single truth from the Speaking Tree, whispered on the wind.
  ‚Ñéfill \textit{The truth is always about a broken promise. The promise may not be yours ‚àí‚àí‚àí but the ninth time you listen, the tree will speak a promise you made and forgot.} \texttt{[TRUTH]}
‚Ñπtem[Q] \textbf{Silk‚àíWarden's Mask} ‚àí‚àí‚àí A carved mask that grants safe passage across any Lethai‚àíar bridge.
  ‚Ñéfill \textit{The mask is also a witness. It will remember every promise you make while wearing it ‚àí‚àí‚àí and the ninth promise will be witnessed by the Hollow.} \texttt{[OATH]}
‚Ñπtem[K] \textbf{Oath‚àíKeeper's Finger} ‚àí‚àí‚àí A dried root shaped like a digit. Use it to call an Oath‚àíKeeper once.
  ‚Ñéfill \textit{The Oath‚àíKeeper will settle one broken promise for you. Then it will ask for yours ‚àí‚àí‚àí and the ninth finger of its hand will point at the promise you most fear to keep.} \texttt{[DEATH]}
‚Ñπtem[A] \textbf{The First Nail} ‚àí‚àí‚àí A legendary iron nail that sealed the first spirit into the Ukwe.
  ‚Ñéfill \textit{Drive it into any tree to create a new Speaking Tree. The tree will remember everything ‚àí‚àí‚àí and the ninth name it speaks will be your own, unless you offer a memory in its place.} \texttt{[LEGEND]}
‚Ñπtem[A+] \textbf{Mami Wata's Tear} ‚àí‚àí‚àí A drop of water that never dries, said to be the spirit's grief for a promise broken long ago.
  ‚Ñéfill \textit{Drink it to forget one broken promise permanently ‚àí‚àí‚àí but you will dream of drowning once each season. The ninth dream will be the promise itself, remembered.} \texttt{[WATER]} \texttt{[MEMORY]}
\end{enumerate}

\section*{Quick use notes}
≈Çabel{sec:sekogo‚àíquick‚àíuse}
\begin{itemize}
‚Ñπtem Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2‚àí‚àí5 $‚Üí$ 4, 6‚àí‚àí10 $‚Üí$ 6, J/Q/K $‚Üí$ 8, A $‚Üí$ 10).
‚Ñπtem Diamonds are codified outcomes (favors, memories, shards) that change position rather than call for a roll.
‚Ñπtem If any A appears, echo \textbf{grove‚àíomens} ‚àí‚àí‚àí bees that fly backward, roots that bleed sap, a mask that watches from the trees, the ninth bell of Ananse's Child ringing in still air. If an A+ appears (special high‚àírank result), a spirit of the deep jungle (Mami Wata, Ananse's Child, the Leopard King) manifests directly.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
≈Çabel{sec:sekogo‚àímechanics}

\begin{itemize}
‚Ñπtem \textbf{The Three Societies:} Sekogo is governed by three overlapping authorities: the Leopard Society (hunters who restore balance), the Earth Society (elders who keep the Calendar of Witness), and the nganga (spirit‚àídoctors who work with \textit{minkisi}). A single elder often belongs to two societies, and the tension between their roles is where drama lives.
‚Ñπtem \textbf{The Ukwe (Speaking Tree):} A living witness whose roots extend across Sekogo. Every broken promise, every forgotten harmony stains its leaves. It answers questions ‚àí‚àí‚àí but never with kindness, and never for free. Accepting its gift creates a reciprocal bond that the tree may call upon at any moment. The ninth root is never still ‚àí‚àí‚àí it searches for the name that was erased.
‚Ñπtem \textbf{The Crimson Valley and Web‚àíLaw:} The Lethai‚àíar guard the valley with fierce neutrality. Their skin‚àítattoos (Marks) record promises witnessed. When a Lethai‚àíar dies, their Marks are cut from the skin and buried at the Ukwe's roots. The web‚àílaw is unwritten; each warden carries a different clause in memory. The ninth clause is always a secret known only to the First Weaver's ghost.
‚Ñπtem \textbf{Mask Law:} In Sekogo, a mask is not a disguise ‚àí‚àí‚àí it is a role. Wearing the mask of a Leopard Society hunter, a Waker, or a spirit entitles you to speak for that office, but also binds you to its taboos. Breaking a mask's taboo (e.g., a Leopard mask showing mercy to a slaver) causes the mask to crack and the wearer to lose $1$ die on all social rolls until they atone. The ninth mask is the Hollow's own ‚àí‚àí‚àí it has no face, and it never cracks because it was never whole.
‚Ñπtem \textbf{Bark Promise‚àíCords (\textit{Mukanda}):} Promises are recorded on strips of bark called \textit{mukanda}. Each strip is knotted; the knot tightens when the promise is strained, loosens when it is kept. A \textit{mukanda} that rots means the bond is released; one that hardens into wood means the obligation has set. The ninth knot on a \textit{mukanda} is never tied ‚àí‚àí‚àí it is left loose for the Hollow to judge.
‚Ñπtem \textbf{Ananse's Bargain (Trickster Mechanic):} Once per session, a character may call upon Ananse's Child to ``renegotiate'' any promise, oath, or contract. Roll Spirit + Lore (DV 4). On a success, the terms shift in your favor ‚àí‚àí‚àí but the GM gains 1 SB and may add a hidden clause that will trigger at the worst possible moment. On a miss, the burden doubles ‚àí‚àí‚àí and the ninth clause added will be the Hollow's own terms.
‚Ñπtem \textbf{The \textit{Mwitu} (Hollow of the Forest):} When the Ninth manifests in Sekogo, it is as the \textit{Mwitu} ‚àí‚àí‚àí the forest that forgets where the village ends. It appears as a wall of gray mist between the trees, and it whispers: ``Why are you still here?'' Those who cannot answer lose their sense of direction (all navigation rolls at $‚àí1$ die for one scene). The \textit{Mwitu} can be appeased by leaving a \textit{mukanda} at the edge of the mist ‚àí‚àí‚àí a promise to remember a forgotten name.
‚Ñπtem \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    ‚Ñπtem \textbf{The Root Father} ‚àí‚àí‚àí Grimmir. \textit{The ancient pact between the jungle and the first honey‚àígatherer. His ninth root is the Hollow's own.}
    ‚Ñπtem \textbf{The Masked One} ‚àí‚àí‚àí Mykkiel. \textit{The law that wears a leopard's face; the judge who hunts. His ninth mask is a blank ‚àí‚àí‚àí for the case he cannot judge.}
    ‚Ñπtem \textbf{The Spider's Whisper} ‚àí‚àí‚àí Inaea. \textit{The trickster who spins webs in the spaces between promises. Her ninth knot is never tied.}
    ‚Ñπtem \textbf{The Water Mother} ‚àí‚àí‚àí Ra√©yn. \textit{Mami Wata, who holds the memories of the drowned. Her ninth tear is the one she never sheds.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross‚àíExpansion Hooks)}
\begin{itemize}
  ‚Ñπtem \textbf{The Ukwe's Shadow and the Hollow (Eastern Realms):} The Hollow is said to be the shadow of the Ukwe ‚àí‚àí‚àí the space where broken promises go when the tree forgets them. Aelaerem hedge‚àíkeepers leave offerings at the Ukwe's roots to keep the Hollow from wandering. The ninth offering is always a \textit{mukanda} with a single knot ‚àí‚àí‚àí a promise to remember.
  ‚Ñπtem \textbf{Lethai‚àíar Silk and the Way of Silk:} The silk bridges of the Crimson Valley are made from spider‚àísap and oath‚àíhair. Tulkani traders prize the silk for memory‚àícords; Fhara water‚àíclerks use it to seal well‚àícontracts. A bolt of Crimson silk can buy safe passage across half the continent ‚àí‚àí‚àí but the ninth bolt is always reserved for the Hollow's bridge.
  ‚Ñπtem \textbf{Leopard Society and Ashaani Slavers:} The Leopard Society has a standing grudge against Ashaani slavers. A masked hunter may offer the party a deal: help free a group of captives from a Black Hand caravan, and the Society will forget an old broken promise. The ninth captive freed will be marked with a Leopard's claw ‚àí‚àí‚àí a sign that they are bound to the Society for a hunt.
  ‚Ñπtem \textbf{Nganga and the Kalrantha Amulet (Dhahara):} The nganga of Sekogo believe the Kalrantha amulet is a rogue \textit{nkisi} ‚àí‚àí‚àí a spirit‚àíobject that was never properly sealed. They want to drive a nail into it to bind the demon. The ninth nail would seal the amulet forever ‚àí‚àí‚àí but the nganga who drives it will become the amulet's new keeper.
  ‚Ñπtem \textbf{Ananse's Child and the Grey Wanderer's Grimoire:} The trickster spirit is known to trade fragments of forgotten prophecies for stories. Some psions claim that Ananse's Child taught the first Echo‚àíReaders the art of reading broken promises in the wind. The ninth prophecy it ever told was about the Hollow's return ‚àí‚àí‚àí but it told it as a joke, and no one laughed.
  ‚Ñπtem \textbf{The Neutral Kingdoms (North):} Smugglers from the unclaimed lands often carry Sekogo honey across the savannah. A Leopard Society hunt that crosses the border may find itself negotiating with a petty king who broke a promise to the Ukwe three generations ago. The king's ninth ancestor was a witness ‚àí‚àí‚àí and his ghost still sits on the throne, waiting for the promise to be kept.
\end{itemize}

\subsection*{Sekogo Superstitions (burn for $+1$ die once)}
\begin{itemize}
  ‚Ñπtem \textbf{Bread to the Root:} Crumble a crust at the Ukwe's base. Ignore the first \emph{Ukwe Demands a Name} penalty. \texttt{[SUPERSTITION]}
  ‚Ñπtem \textbf{Knot the Silk:} Tie a single knot in a Lethai‚àíar bridge cord. Cancel \emph{Web‚àíLaw Challenge} or take $‚àí1$ die on your next negotiation (your choice). \texttt{[SUPERSTITION]}
  ‚Ñπtem \textbf{Name to the Nail:} Whisper a dead nganga's name into a \textit{nkisi}. Gain $+1$ Effect when using a spirit‚àíobject this scene, but the ancestors will expect an answer ‚àí‚àí‚àí and the ninth answer must be a promise. \texttt{[SUPERSTITION]}
  ‚Ñπtem \textbf{Spit to the Mask:} Before donning a ceremonial mask, spit on the inside. If the mask accepts, you gain $+1$ die to all rolls while wearing it. If it rejects, the mask will not come off for $1\text{d}4$ hours ‚àí‚àí‚àí and the ninth hour, the mask will speak. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
≈Çabel{sec:sekogo‚àísb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
‚Ñπtem \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Leopard Society hunters or Earth Society elders}.
‚Ñπtem \textbf{Noise:} Sounds of your actions alert nearby \textbf{bees or Lethai‚àíar wardens}.
‚Ñπtem \textbf{Trace:} Evidence of your passage marks your route for \textbf{Oath‚àíKeepers or Waker spies}.
‚Ñπtem \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or safe passage through a grove}.
‚Ñπtem \textbf{Supply Strain:} Mark $+1$ segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
‚Ñπtem \textbf{Alarm Raised:} \textbf{Local Leopard Society or Green Council druid} becomes aware and begins responding.
‚Ñπtem \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{thorn‚àíspirit or fog from the Crimson Falls}.
‚Ñπtem \textbf{Foe Appears:} A \textbf{Waker cell or Oath‚àíKeeper} arrives on scene.
‚Ñπtem \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
‚Ñπtem \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
‚Ñπtem \textbf{Reinforcements:} Additional \textbf{Leopard Society, Earth Society guardians, or Lethai‚àíar archers} arrive.
‚Ñπtem \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
‚Ñπtem \textbf{Major Twist:} The situation fundamentally changes ‚àí‚àí‚àí \textbf{Ukwe speaks/Waker poison spreads/web‚àílaw broken}.
‚Ñπtem \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
‚Ñπtem \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
‚Ñπtem \textbf{Trap Springs:} A prepared danger activates with full effect.
‚Ñπtem \textbf{Authority Arrival:} \textbf{Leopard King, Ukwe's Voice, or Crimson Valley First Weaver} intervenes.
‚Ñπtem \textbf{Scene Shift:} The environment changes dramatically ‚àí‚àí‚àí \textbf{\textit{nkisi} shatter/honey turns to blood/Ukwe's root breaches the ground}.
‚Ñπtem \textbf{Patron Omen:} Divine/arcane forces take notice ‚àí‚àí‚àí \textbf{omen appears/blessing lost/curse manifests}.
‚Ñπtem \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region‚àíSpecific SB Options}
\begin{itemize}
‚Ñπtem \textbf{Sekogo (Jungle Law):} Bees form words, tree bark weeps sap, the Ukwe's root trembles.
‚Ñπtem \textbf{Sekogo (Crimson Valley):} Silk cords tighten, canyon walls bleed rust, the web‚àílaw recites itself.
‚Ñπtem \textbf{Sekogo (Society Conflict):} Masks shift expressions, bark strips burn without flame, a ninth nail appears in every \textit{nkisi}.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
‚Ñπtem \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Thorn Wall}.
‚Ñπtem \textbf{Cut the Web (Presence + Sway):} On a hit, freeze range for one exchange; if a silk cord is in your possession, $+1$ die.
‚Ñπtem \textbf{Weather the Hunt (Resolve + Spirit):} On a hit in \emph{Leopard Society Hunt}, you pick who is targeted; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Root & Mask)}
\begin{itemize}
‚Ñπtem \textbf{The Root‚àíKing (Nine Broken Promises)} ‚àí‚àí‚àí \emph{Phases}: Promise Calls $‚Üí$ Orchestrated Famine $‚Üí$ Grove Coup. \emph{Cracks}: burn a bark strip, expose a false memory, survive a harvest without honey. \textit{The ninth root follows you home.}
‚Ñπtem \textbf{The Mask that Walks (Leopard's Shadow)} ‚àí‚àí‚àí \emph{Phases}: Midnight Hunts $‚Üí$ Procession of Masks $‚Üí$ Skin‚àíBaptism of a Warden. \emph{Cracks}: relic leopard claw, survivor testimony, the mask turning against them ‚àí‚àí‚àí and the ninth mask is always the Hollow's own. 
‚Ñπtem \textbf{Mami Wata's Wrath (Water Spirit)} ‚àí‚àí‚àí \emph{Phases}: Pool Rises $‚Üí$ Crimson Falls Reversed $‚Üí$ Memory of the Drowned. \emph{Cracks}: offer a true memory of loss, speak the name of a slaver ship she sank, return a stolen mirror to her pool. \textit{The ninth memory she takes is your own.}
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage & Power}]
Among Sekogo, power flows through witnessed promises, honey, and the consent of the jungle. The Green Council rules by consensus, but the Leopard Society restores balance, the Earth Society keeps the Calendar of Witness, and the nganga negotiate with spirits. No single faction holds all the power ‚àí‚àí‚àí and that is by design.

\textbf{For the GM:}  
Power in Sekogo society turns on broken promises and witnessed oaths. Patronage often takes the form of favors, memories, or shards ‚àí‚àí‚àí each carrying a future cost. To emphasize this:
\begin{itemize}
‚Ñπtem Tie rewards to objects that create reciprocal bonds (leopard claws, Ukwe leaves, \textit{nkisi} shards, \textit{mukanda} strips) that can be called in at inconvenient moments.
‚Ñπtem Let rival societies offer conflicting judgments, forcing players to choose whose law to follow ‚àí‚àí‚àí and whose ninth clause to honor.
‚Ñπtem Use the honey‚àíorchards, druid circles, and silk bridges as arenas for social contests, where a gift of honey is a treaty and a broken promise is a death sentence.
‚Ñπtem Introduce trickster spirits (Ananse's Child) to complicate simple resolutions ‚àí‚àí‚àí they are excellent wild cards, and their ninth bell is always the Hollow's cue.
‚Ñπtem Remember that Mami Wata is not evil; she is ancient. She collects memories, not lives. A bargain with her may be the only way to escape a blood‚àífeud ‚àí‚àí‚àí but the ninth memory she asks for will be the one you cannot spare.
\end{itemize}
In Sekogo society, the jungle remembers everything ‚àí‚àí‚àí and the Ukwe forgets nothing. The \textit{Mwitu} waits at the edge of the ninth root, and the ninth bell of Ananse's Child rings only when a promise is already broken.
\end{tcolorbox}

% Index entries
‚ààdex{Sekogo|see{Green Council}}
‚ààdex{Theme generator}
‚ààdex{Green Council generator}
‚ààdex{Places!Sekogo}
‚ààdex{People!Sekogo}
‚ààdex{Complications!Sekogo}
‚ààdex{Rewards!Sekogo}
‚ààdex{Leopard Society}
‚ààdex{Earth Society}
‚ààdex{Nganga}
‚ààdex{Ukwe (Speaking Tree)}
‚ààdex{Crimson Valley}
‚ààdex{Lethai‚àíar}
‚ààdex{Web‚àílaw}
‚ààdex{Marks (tattoos)}
‚ààdex{Honey‚àíorchards}
‚ààdex{Minkisi (power objects)}
‚ààdex{Wakers vs Conservators}
‚ààdex{Ananse (trickster)}
‚ààdex{Mami Wata}
‚ààdex{Ninki Nanka}
‚ààdex{Mwitu (Hollow of the forest)}
‚ààdex{Mukanda (promise‚àícords)}
‚ààdex{Local Patrons!Sekogo}