\section{Sidhi Homeland −−− Galanina, ``The Sealed Coast''}
łabel{chap:galanina}

\textbf{Key Demographic: } Human
\textbf{Capital: } Galanina (city−state)

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The sea remembers every child. The green flame burns for the unnumbered.}

\begin{quote}
  \textbf{High Keeper of the Unforgotten (Elite):}  
  ``The Ashaani came with brass chains and hungry gods. They took our children −− not to kill, but to \textit{unname}. The Covenant drove them back, but the children did not return. We light green flames on every tower so the lost may find their way home. The flames have never stopped.''
\end{quote}

\begin{quote}
  \textbf{Lighthouse Freedman (Commoner):}  
  ``My grandmother was a child when the Ashaani took her. She came back −− but not whole. She forgot her mother's face. She remembered only the scent of myrrh and the sound of a bell that rang nine times. When she died, we lit a green lamp for her. It still burns. It will burn until she remembers.''
\end{quote}

\textbf{Genre / Mood:} Levantine coastal grief, covenant justice, the weight of unnumbered children, green flame memorials.

\textbf{Starting Location:} The Great Lighthouse of Galanina, where a green flame burns in a cage of brass. A High Keeper sits on a stone bench, a list of names longer than her arm. ``The Covenant says the Ashaani are gone. But every spring, a child walks into the sea. We find their shoes on the shore. No footprints leading away. The sea is not the sea −− it is a door. And someone is leaving it open.''

\vspace{2mm}
\begin{center}
łine(1,0){250}
\end{center}
\textbf{Theme & Atmosphere}

Galanina is a limestone city on a promontory, its white walls streaked with salt and grief. The Ashaani occupied it for a generation, taking children to be hollowed into spirit vessels. The Covenant broke the siege, but the children never fully returned −− some came back empty, their eyes the colour of old brass. Now Galanina is a free city under Covenant protection, a waystation for the Sidhi diaspora. Green flames burn in every tower, every home, every lighthouse. The living tend the flames for the unnumbered dead. The sea glows green on winter solstice. And every spring, a child walks into the waves. No one knows why. The Hollow watches from the Unnumbered Shore.

\textbf{What you notice first:} The smell of olive oil and sea salt. The soft ring of brass bells on every doorway. The way every Sidhi touches their lamp before speaking a child's name. The silence when the green flame flickers.

\textbf{The one rule that kills newcomers:} Never extinguish a green lamp. The flame remembers the child it was lit for. To kill the flame is to kill the memory −− and the Hollow will ask why.

\subsection*{The Unnumbered List (Campaign Timer)}
The children taken by the Ashaani are called the Unnumbered −− their names were never recorded, or were erased. Track the Hollow's interest in the unnumbered with a visible timer:

\textbf{Unnumbered Grief [8]} – Advances when:
\begin{itemize}
ℹtem A character extinguishes a green lamp (+1)
ℹtem A child disappears from the shore (+2)
ℹtem A Hollow Child speaks (GM spends 1 SB to tick +1)
ℹtem The Ashaani bell rings nine times (+2)
ℹtem The party learns a name from the Sealed Archive (+1)
\end{itemize}
At 4 segments: Green flames burn low. The sea is still. The Hollow Children turn their heads toward the shore.  
At 8 segments: The Unnumbered Shore opens. Roll on the \textbf{Hollow Walks} table. The party must offer a name (living or dead) or the Hollow takes one.

\subsection*{Hollow Walks (d6)}
When the Unnumbered Grief timer reaches 8, roll 1d6 for the manifestation:
\begin{enumerate}
ℹtem A ghost child appears on the shore. It asks: ``Where is my mother?'' The first PC to answer must give a true name of a living person. The ghost will follow that person for a day.
ℹtem The sea throws back a child's shoe. Inside is a note: ``We are still waiting.'' The next green lamp the party touches gutters. Tick the timer back to 0, but the GM gains 3 SB.
ℹtem A Hollow Child points at a PC. That PC now hears children's laughter in every empty room. Until they make an offering at the Ninth Name Shrine, all stealth rolls are Desperate.
ℹtem The Oath−Keeper appears at the Green Bazaar. It asks for a name from the Sealed Archive. If the party gives one, the timer resets. If they refuse, the Hollow takes the name of the nearest child (NPC) – that child vanishes.
ℹtem The Great Lighthouse flame turns blue. The High Keeper weeps. For the next scene, any lie spoken within Galanina is heard by every green lamp in the city. The GM gains 1 SB per lie.
ℹtem The Unnumbered Shore is covered with wet footprints – hundreds of them, all leading inland. The Hollow Children walk among the living for one night. They do not speak. They only watch. Any PC who fails a Resolve test (DV 4) forgets the face of a loved one for a day.
\end{enumerate}

\subsection*{Lamp Etiquette Violations (d6)}
When a PC breaks lamp custom (extinguishes a green lamp, refuses to light one for a child, steals a lamp's name), roll 1d6:
\begin{center}
\begin{tabular}{|c|l|}
ℎline
\textbf{d6} & \textbf{Consequence} \\
ℎline
1 & The lamp's child appears in a dream. It asks for a story. If you cannot tell one, you wake with 1 Fatigue. \\
2 & The green flame spits ash at you. You gain the Condition \textit{Ash−Touched} (–1 die on all social rolls with Sidhi for a week). \\
3 & A Hollow Child follows you. It does not attack. It only stares. Until you make amends, all stealth rolls are Desperate. \\
4 & The Oath−Keeper writes your name in its ledger. The Hollow will visit you within nine days unless you light a lamp for an unnumbered child (costs 1 Supply). \\
5 & The High Keeper bans you from the Great Lighthouse. You cannot access the Sealed Archive or the Unnumbered Shore. The ban lifts only after you recover a stolen lamp. \\
6 & The sea rises. A wave washes over your camp. Lose 1 Supply (salt damage) and tick \textbf{Unnumbered Grief +1}. The Hollow Children laugh. \\
ℎline
\end{tabular}
\end{center}

\subsection*{Prominent NPCs of Galanina}
łabel{sec:galanina−npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
ℎline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
ℎline
High Keeper Eliron & Keeper of the Great Lighthouse & Remember the unnumbered, one name at a time & She has nine hundred names. She does not know the last nine. \\
ℎline
Lamp−Maker Shalvi & Artisan of brass and glass & Make a lamp that never gutters & Her hands are stained green. She cannot wash it off. \\
ℎline
Covenant Judge Vered & Silent arbiter of the Sealed Gate & Keep the names from the Ashaani & She has not spoken in nine years. Her judgments are written. \\
ℎline
The Hollow Child (unnamed) & One of the returned children & Watch. Always watch. & Her eyes are brass. She never blinks. \\
ℎline
Covenant Marine Captain Yaron & Commander of the \textit{Reckoning} & Hunt the slavers who took his daughter & He carries her green lamp. He has never lit it. \\
ℎline
Ashaani Remnant (spy) & Veiled Sister in disguise & Recover the names from the Sealed Archive & Her brass bell rings when she lies. It rings often. \\
ℎline
\end{longtable}

\subsection*{Names of Galanina}
łabel{sec:galanina−names}

Levantine / Phoenician / Canaanite influences: soft consonants, vowel endings, familial suffixes.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
ℎline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surname / House} & \textbf{Epithet} \\
ℎline
Eliron & Tamira & Gubla−im & \textit{the Unforgotten} \\
ℎline
Yaron & Shalvi & Byblos−on & \textit{the Lamp−Born} \\
ℎline
Adoni & Elissa & Sidoni & \textit{the Green−Flamed} \\
ℎline
Zuriel & Ba'alat & Tyri−ba'al & \textit{the Unnumbered} \\
ℎline
Mahir & Anat & Arvad−im & \textit{the Hollow−Watched} \\
ℎline
Shemarya & Hodaya & Zephon−on & \textit{the Ninth Child} \\
ℎline
\end{longtable}

\paragraph*{(Lighthouse/Shore/Quarter)} Great Lighthouse with green flame; Unnumbered Shore with wet sand and children's footprints; Lamp−Maker's Quarter with brass and glass.

\section*{Spades −−− Places}
łabel{sec:galanina−places}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem \textbf{Great Lighthouse of Galanina} ℎfill \textit{[LIGHT] [DEATH]}  
  Limestone tower, green flame visible for a day's sail. The oil is from olives grown in the Vale of Tears. The lighthouse keeper is blind; she saw the ninth child walk into the sea.  
  \textbf{Mechanical hook:} Lighting a lamp here (Spirit+Lore DV 3) clears 1 Fatigue but gives you a vision of a missing child. The GM gains 1 SB.

ℹtem \textbf{The Unnumbered Shore} ℎfill \textit{[THRESHOLD] [DEATH]}  
  Beach where green lamps are placed each solstice. At low tide, remains of an Ashaani pier. On moonless nights, children's footprints appear. They lead to the water. They never lead back.  
  \textbf{Mechanical hook:} A character who follows the footprints must test Resolve (DV 5). On failure, they walk into the sea −− rescued only if another PC acts immediately.

ℹtem \textbf{Covenant Citadel (Sealed Gate)} ℎfill \textit{[JUSTICE] [SECRET]}  
  Grey fortress on Ashaani temple ruins. Walls carved with names of taken children. The ninth name is blank. The Covenant judge has not spoken in nine years.  
  \textbf{Mechanical hook:} Asking the judge a question requires a written petition. The judge's answer is always a single word. That word is a name.

ℹtem \textbf{Lamp−Maker's Quarter} ℎfill \textit{[CRAFT] [MEMORY]}  
  Narrow alleys where artisans hammer brass and blow glass. Oldest lamps have names etched inside. The lamps hum at midnight.  
  \textbf{Mechanical hook:} Buying a lamp (1 Supply) gives you a \textbf{Green Lamp} (see Diamonds). The lamp−maker will ask for a name to etch. If you give a false name, the lamp gutters.

ℹtem \textbf{The Sealed Archive} ℎfill \textit{[SECRET] [WEIGHT]}  
  Subterranean vault beneath the lighthouse. Spiral stair of nine steps. The ninth step is a trap. Inside: the Covenants' copies of Ashaani treaties. The ninth treaty lists the children given away. The ink is still wet.  
  \textbf{Mechanical hook:} Reading the ninth treaty (Wits+Lore DV 5) reveals one name of a taken child. Tick \textbf{Unnumbered Grief +1}. The name can be used to bargain with the Hollow.

ℹtem \textbf{The Green Bazaar} ℎfill \textit{[TRADE] [DEATH]}  
  Market where everything is sold by lamp−light. Real trade is in names. A forgotten name can be bought for a silver coin. The Hollow visits on the ninth night of every month.  
  \textbf{Mechanical hook:} Buying a name (1 Boon) gives you a \textbf{Forgotten Name} (once, use to compel a confession from anyone who knew that person). The Hollow will remember you.

ℹtem \textbf{The Hollow Children's Home} ℎfill \textit{[FEAR] [MYSTERY]}  
  Stone building with no windows, set back from the shore. The hollowed children live here. They do not speak, eat, or sleep. They only watch. Their eyes are the colour of old brass.  
  \textbf{Mechanical hook:} Spending a scene watching with them (no roll) gives you a vision of the Ashaani ritual. The GM must answer one question about the Hollow Children's origin.

ℹtem \textbf{Vale of Tears} ℎfill \textit{[RITUAL] [SORROW]}  
  Olive grove where the Ashaani camped. The olives are bitter. Oil from them burns green. Women weep here for the unnumbered. Their tears are collected and poured into the lighthouse flame.  
  \textbf{Mechanical hook:} Collecting a tear (Wits+Survival DV 3) gives you a \textbf{Tear of the Vale} (see Diamonds). Drinking a tear restores 1 Fatigue but gives you a vision of a child's disappearance.

ℹtem \textbf{The Ninth Name Shrine} ℎfill \textit{[DEATH] [FEAR]}  
  Stone chapel at the edge of the Unnumbered Shore. Brass plaque with nine blank spaces. Pilgrims scratch their own names into the brass. The names vanish by morning. The Hollow collects them.  
  \textbf{Mechanical hook:} Scratching your name gives the GM a permanent String. You may ask one question of the Hollow; it will answer truthfully but cryptically.

ℹtem[J] \textbf{The Covenant Docks} ℎfill \textit{[AUTHORITY] [OMEN]}  
  Fortified harbour where Covenant warships anchor. Ships are white with green trim. Ninth ship in every fleet is named \textit{Unnumbered}. It has never returned from patrol.  
  \textbf{Mechanical hook:} A captain offers you passage. The price is a name of a child you knew. If you cannot give one, you must sail on the \textit{Unnumbered}. The ship will return −− but you may not.

ℹtem[Q] \textbf{The Brass Bell Tower} ℎfill \textit{[OMEN] [DEATH]}  
  A tower with nine brass bells. The ninth bell is cracked. It rings only when a child is about to be taken. The sound is a scream. No one climbs the tower.  
  \textbf{Mechanical hook:} If the ninth bell rings during a scene, tick \textbf{Unnumbered Grief +2}. The party has one exchange to find the child before the Hollow takes them.

ℹtem[K] \textbf{The Ashaani Well} ℎfill \textit{[WATER] [CURSE]}  
  A stone well in the old quarter, sealed by the Covenant. The water is black. It tastes of myrrh and copper. The Ashaani used it to wash the children before the ritual. The well still whispers.  
  \textbf{Mechanical hook:} Drinking from the well (Resolve test DV 4) gives you a vision of the ritual. On failure, you gain the Condition \textit{Well−Touched} (nightmares; –1 die on Resolve tests) for a week.

ℹtem[A] \textbf{The Unnamed Gate (Myth)} ℎfill \textit{[LEGEND] [THRESHOLD]}  
  A door that appears only on the winter solstice, on the Unnumbered Shore. It is made of brass and bone. Open it, and you may walk into the Hollow's domain. The price is a name you have never spoken.  
  \textbf{Mechanical hook:} Entering the gate resets \textbf{Unnumbered Grief} to 0 but gives the GM 5 SB. The party emerges one day later with no memory of what they saw −− except that they know one name from the Sealed Archive.
\end{enumerate}

\paragraph*{(Keepers/Makers/Children)} High Keeper with a list of names; lamp−maker with green−stained hands; a Hollow Child who watches.

\section*{Hearts −−− People & Factions}
łabel{sec:galanina−people}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem \textbf{High Keeper Eliron} ℎfill \textit{[AUTHORITY] [SORROW]}  
  Aged, one eye missing (taken by an Ashaani interrogator). She tends the Great Lighthouse. She whispers the names of the unnumbered each dusk. She has nine hundred names. She has nine more to learn before she dies. She does not know whose they are.  
  \textbf{What they carry:} She knows the location of the Sealed Archive's ninth lock. She will trade the secret for a name of a living child (to add to her list).  
  \textbf{What they ask:} A day of silent vigil at the lighthouse (no roll). If you speak, you must leave.

ℹtem \textbf{Lamp−Maker Shalvi} ℎfill \textit{[CRAFT] [MEMORY]}  
  Woman with brass−stained hands and a copper bell on her belt. She makes lamps for the unnumbered. Each lamp is a prayer. She has made nine hundred lamps. The nine hundredth lamp lit itself when she finished it. It shows a face she does not recognise – her own, as a child, before the Ashaani took her sister.  
  \textbf{What they carry:} She will make you a lamp (1 Supply). If you give her a name to etch, the lamp will never gutter. The name must be true.  
  \textbf{What they ask:} A story about a child who was lost. She knots it into the lamp's wick.

ℹtem \textbf{Covenant Judge Vered} ℎfill \textit{[JUSTICE] [SECRET]}  
  Silent figure in grey, carrying a ledger bound in green leather. She never speaks. She writes her judgments on clay tablets. The tablets are fired in the Great Lighthouse flame. The unnumbered can read them. The living cannot.  
  \textbf{What they carry:} She has a tablet with the names of the nine children the Covenant traded. She will show it to anyone who brings her a lamp stolen from a Hollow Child.  
  \textbf{What they ask:} A written confession of a crime you have committed. She will not say why.

ℹtem \textbf{The Hollow Child (unnamed)} ℎfill \textit{[FEAR] [OMEN]}  
  One of the returned children. She has not aged in fifty years. Her eyes are brass. She never blinks. She sits on the shore at low tide, facing the sea. She will not speak – but if you sit beside her, she will point at the horizon. The direction she points is where the next child will disappear.  
  \textbf{What they carry:} Nothing. She does not trade. She only witnesses.  
  \textbf{What they ask:} Your name. If you give it, she will remember it. The Hollow will know it.

ℹtem \textbf{Covenant Marine Captain Yaron} ℎfill \textit{[COMBAT] [GRIEF]}  
  Scarred veteran who lost his daughter to the Ashaani. He hunts slavers now. His ship is named \textit{Reckoning}. He carries a green lamp on his belt – his daughter's. He has never lit it. He is waiting for her to come home.  
  \textbf{What they carry:} He will escort you through Covenant waters for free if you promise to help him find a slaver ship. The promise is a lamp oath – break it, and your own lamp will gutter.  
  \textbf{What they ask:} A name of a child who was taken but not listed. He adds it to his daughter's lamp.

ℹtem \textbf{Ashaani Remnant (Spy)} ℎfill \textit{[DECEPTION] [POLITICS]}  
  Veiled Sister in disguise, posing as a Dhaharan spice merchant. She seeks the names of the children in the Sealed Archive – to finish the ritual her ancestors began. She carries a brass bell that rings only when she lies. It rings often.  
  \textbf{What they carry:} She offers forged Covenant writs (1 Boon) that let you bypass checkpoints. The writs are stamped with a bell. The bell's ring is false.  
  \textbf{What they ask:} Access to the Sealed Archive. She will pay in Ashaani gold – cursed, but still spends.

ℹtem \textbf{Olive−Grove Keeper Tamira} ℎfill \textit{[RITUAL] [MYSTERY]}  
  Woman who tends the Vale of Tears. She collects tears in clay jars. She knows whose tears are whose. She will not tell. She says the tears are a weight that must be carried. The Hollow visits her each new moon. They talk. She does not say about what.  
  \textbf{What they carry:} A jar of tears (1 Supply). Pour them into any lamp, and the flame will show the face of the person who wept them.  
  \textbf{What they ask:} A tear of your own. She will not explain.

ℹtem \textbf{Sidhi Freedman Merchant} ℎfill \textit{[TRADE] [HERITAGE]}  
  Trader from the Brass Coast diaspora, returned to Galanina to honour his grandmother's lamp. He sells Galaninan glass in Dhahara and Dhaharan spice in Galanina. He carries two lamps – one green (for the unnumbered), one brass (for profit). He never mixes the flames.  
  \textbf{What they carry:} He offers a \textbf{Green Lamp} (see Diamonds) for 1 Supply or a story about a child who returned.  
  \textbf{What they ask:} A name to carve into a lamp. He will not say whose.

ℹtem[J] \textbf{The Ninth Child's Ghost} ℎfill \textit{[DEATH] [LEGEND]}  
  Translucent figure seen only at the Ninth Name Shrine. She was the last child the Ashaani took. She did not go quietly. She clawed the name ``Covenant'' into the brass before she vanished. The scratch marks are still there. They bleed rust.  
  \textbf{What they carry:} She will give you a shard of brass from the plaque. The shard is warm. It can be used as a witness for any oath (once).  
  \textbf{What they ask:} A promise that you will remember her name. Her name is scratched on the shard. It is the only copy.

ℹtem[Q] \textbf{The Hollow's Accountant (Oath−Keeper)} ℎfill \textit{[DEATH] [WEIGHT]}  
  Faceless figure seen at the Green Bazaar on the ninth night. It asks only one question: ``What name did you forget?'' If you answer truthfully, it writes the name in its ledger and leaves. If you lie, it writes your name instead. The Hollow will visit you within nine days.  
  \textbf{What they carry:} It cannot be bargained with. Only a true name of a taken child can satisfy it.  
  \textbf{What they ask:} A name. It already knows.

ℹtem[K] \textbf{The First Freedman (Ghost)} ℎfill \textit{[LEGEND] [JUSTICE]}  
  The first Sidhi to buy his freedom after the Ashaani occupation. He sits at the bottom of the Ninth Name Shrine. He asks every visitor: ``Did you light a lamp for me?'' No one knows his name. He will not tell.  
  \textbf{What they carry:} If you light a lamp for him (1 Supply), he will teach you a \textbf{Lamp Oath} – a promise that cannot be broken without the Hollow's attention.  
  \textbf{What they ask:} A name. He collects them. He will not say why.

ℹtem[A] \textbf{The Unnumbered Mother (Myth)} ℎfill \textit{[PATRON] [SORROW]}  
  A figure in grey who walks the Unnumbered Shore at dawn. She was the mother of the first child taken. She walks still, calling names that no one answers. If you hear her call your name, you have one day to find a child to take your place – or you will walk into the sea.  
  \textbf{What they carry:} She offers no bargains. She only collects.  
  \textbf{What they ask:} Your name. She will not stop calling until you give it.
\end{enumerate}

\paragraph*{(Flame/Shore/Bell)} Green flame gutters – a child's lamp goes out; the Unnumbered Shore weeps – sea throws back a child's shoe; the ninth bell rings – a child is about to vanish.

\section*{Clubs −−− Complications/Threats}
łabel{sec:galanina−complications}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem \textbf{Lamp Gutters} ℎfill \textit{[DEATH] [WEIGHT]}  
  A child's memorial lamp flickers and dies. The flame cannot be relit. The child's name is now unnumbered. The next night, that child walks into the sea. The party can prevent it only by giving the Hollow a new name – a living one.  
  \textbf{Mechanical hook:} Tick \textbf{Unnumbered Grief +1}. The party has until midnight to find a substitute name or a story that satisfies the Hollow.

ℹtem \textbf{Covenant Judgment} ℎfill \textit{[JUSTICE] [FEAR]}  
  The silent judge writes a name on a clay tablet. The name is yours. You have nine days to prove you are not a slaver – or to flee Galanina. The Covenant does not accept appeals. The green flame will witness your trial.  
  \textbf{Mechanical hook:} A PC must appear before the judge. They may present evidence (Wits+Investigation DV 4) or a witness (Presence+Sway DV 4). Failure results in exile – the PC cannot return to Galanina for a year.

ℹtem \textbf{Ashaani Bell Rings} ℎfill \textit{[COMBAT] [DECEPTION]}  
  A Veiled Sister's bell rings nine times. The Hollow Children stir. They point at the sea. A ship is coming – a slaver ship disguised as a Dhaharan trader. The party has one hour to warn the Covenant or to stop the ship themselves.  
  \textbf{Mechanical hook:} The ship has Cap 4 slavers. If the party stops it, gain a \textbf{Covenant Writ of Passage}. If they fail, tick \textbf{Unnumbered Grief +2} and a child is taken.

ℹtem \textbf{The Unnumbered Shore Weeps} ℎfill \textit{[OMEN] [DEATH]}  
  The sea throws back a child's shoe. It is still warm. Inside is a note: ``We are still waiting.'' The writing is in Ashaani. The shoe belonged to a child taken fifty years ago. The next night, the Hollow walks the shore.  
  \textbf{Mechanical hook:} The party must make an offering (1 Supply or a story) at the Ninth Name Shrine before dawn. Failure means the Hollow takes a name from the party (GM chooses one Bond to erase temporarily).

ℹtem \textbf{Lamp−Maker's Bargain} ℎfill \textit{[TRADE] [SORROW]}  
  Shalvi offers to make you a lamp that never goes out. The price is a memory of a child you loved. If you accept, you forget that child completely. The lamp burns green. The child's name is added to the Unnumbered List.  
  \textbf{Mechanical hook:} The player loses one Bond (the child). The lamp acts as a \textbf{Green Lamp} (see Diamonds) but cannot be used again. The Hollow will remind you of the child's name in dreams (GM gains 1 SB per session).

ℹtem \textbf{The Ninth Name Fills} ℎfill \textit{[DEATH] [FEAR]}  
  The ninth blank on the brass plaque suddenly shows a name: yours. The Hollow Children turn their heads toward you. The Oath−Keeper appears. It asks: ``When did you forget?'' You have nine breaths to answer. If you cannot, the Hollow takes your shadow.  
  \textbf{Mechanical hook:} The PC must answer truthfully (Resolve test DV 5). On failure, they lose their shadow for a week (–1 die on all stealth rolls). The GM gains a permanent String on them.

ℹtem[J] \textbf{Green Bazaar Closed} ℎfill \textit{[TRADE] [OMEN]}  
  The market is empty. Stalls are shuttered. Lamps are dark. The Hollow visited last night. It took nine names. The merchants are afraid. They will not trade until a new lamp is lit for each name.  
  \textbf{Mechanical hook:} The party must light nine lamps (1 Supply each) or find another source of supplies. Until then, all trade in Galanina is at +1 DV.

ℹtem[Q] \textbf{Hollow Child Speaks} ℎfill \textit{[FEAR] [MYSTERY]}  
  One of the Hollow Children opens its mouth. It speaks a single word: a name – the name of a PC. That PC now sees the Hollow Children everywhere. They do not attack. They only watch.  
  \textbf{Mechanical hook:} The PC suffers –1 die on all Resolve tests until they make an offering at the Great Lighthouse (1 Supply). The Hollow Child will not speak again.

ℹtem[K] \textbf{Covenant Fleet Blockade} ℎfill \textit{[AUTHORITY] [WAR]}  
  The Covenant has sealed the harbour. No one leaves. A slaver ship escaped last week with a child. The Covenant blames the Sidhi. Tensions are high. The party is caught in the middle.  
  \textbf{Mechanical hook:} The party must prove the Sidhi are innocent (Investigation test DV 4) or help the Covenant find the real slavers. Failure leads to a riot – treat as a Cap 3 mob.

ℹtem[A] \textbf{The Unnamed Gate Opens} ℎfill \textit{[LEGEND] [APOCALYPSE]}  
  The brass−and−bone door appears on the Unnumbered Shore. It is not solstice. It should not be here. The sea is calm. Too calm. The Hollow Children walk toward it, one by one. If they all enter, the Hollow will have nine new servants.  
  \textbf{Mechanical hook:} The party has one scene to stop the children or to enter the gate themselves. Tick \textbf{Unnumbered Grief} to maximum. The Hollow now walks openly in Galanina.
\end{enumerate}

\paragraph*{(Lamp/Name/Tear)} Green lamp (once, calm the Hollow); forgotten name (compel a confession); tear of the Vale (restore Fatigue, gain a vision).

\section*{Diamonds −−− Treasures & Discoveries}
łabel{sec:galanina−rewards}
\begin{enumerate}
\setcounter{enumi}{1}
ℹtem \textbf{Green Lamp (Memorial)} ℎfill \textit{[PROTECTION] [ONCE]}  
  Small brass lamp that burns green without oil. Carry it to calm the Hollow for one scene. The flame flickers when a child is about to be taken – you have one exchange to prevent it. The lamp cannot be used again.  
  \textbf{Hollow cost:} After use, the lamp's wick turns black. The Hollow knows you prevented a taking. It will remember.

ℹtem \textbf{Lamp−Maker's Name} ℎfill \textit{[SOCIAL] [TRUTH]}  
  Brass token with a name scratched on it. Present it to any Galaninan to demand a single truth. The name is of a child who was taken. The speaker cannot lie – the child's ghost is watching. The token crumbles after use.  
  \textbf{Hollow cost:} The ghost of that child will visit you in dreams until you light a lamp for them (1 Supply).

ℹtem \textbf{Sealed Archive Key} ℎfill \textit{[SECRET] [WEIGHT]}  
  Brass key that opens the ninth lock of the Sealed Archive. The ninth lock is a bell. Turn the key, and the bell rings once. The archive door opens – but the Oath−Keeper will know you entered. You have nine minutes inside.  
  \textbf{Hollow cost:} The Oath−Keeper will ask you one question about what you saw. If you lie, it writes your name in its ledger.

ℹtem \textbf{Covenant Writ of Passage} ℎfill \textit{[AUTHORITY] [TRAVEL]}  
  Clay tablet stamped with the Covenant seal. Safe passage through any Covenant checkpoint. The seal is warm. If you break Covenant law, the seal burns your skin (1 Harm). The writ is valid for one journey.  
  \textbf{Hollow cost:} The seal's warmth attracts the Hollow. Tick \textbf{Unnumbered Grief +1} each time you use it.

ℹtem \textbf{Hollow Child's Pointing} ℎfill \textit{[OMEN] [DEBT]}  
  A child points at a person. That person will be the next to vanish – or will be the one who saves the next child. The GM must reveal which. The pointing can be used once per session. The child does not speak. They never lie.  
  \textbf{Hollow cost:} After using, the child follows you for a day. The GM may spend 1 SB to have them point again – at a PC.

ℹtem \textbf{Tear of the Vale} ℎfill \textit{[RITUAL] [PROTECTION]}  
  Vial of tears collected from the olive grove. Pour it into any water, and the Hollow will avoid crossing that water for one day. The water becomes bitter. Drinking it restores 1 Fatigue but gives you a vision of a child who was taken. The vision costs 1 Corruption.  
  \textbf{Hollow cost:} The Hollow will remember your scent. The next time you cross water, the GM may spend 1 SB to have the Hollow appear.

ℹtem[J] \textbf{Brass Shard (Ninth Child's Plaque)} ℎfill \textit{[WITNESS] [ONCE]}  
  Shard of brass from the Ninth Name Shrine. Touch it to any oath to make it witnessed by the unnumbered. Breaking the oath ticks \textbf{Unnumbered Grief +2} and the Hollow will visit the oath−breaker. The shard crumbles after use.  
  \textbf{Hollow cost:} The ninth child's ghost will appear and ask: ``Did you keep your word?'' If you broke the oath, she takes a memory.

ℹtem[Q] \textbf{Lamp Oath (The First Freedman's Teaching)} ℎfill \textit{[VOW] [ONCE]}  
  A promise sworn over a green lamp. The oath is binding – breaking it extinguishes the lamp and ticks \textbf{Unnumbered Grief +3}. The oath can be used as a String against the oath−taker.  
  \textbf{Hollow cost:} The lamp's flame turns blue while the oath is active. The Hollow can see you through the flame.

ℹtem[K] \textbf{Unnumbered List (Page)} ℎfill \textit{[WEIGHT] [SECRET]}  
  A single page from the High Keeper's list. It contains nine names of taken children. Speak a name aloud, and the Hollow will answer one question about that child. The page burns after nine uses.  
  \textbf{Hollow cost:} Each time you use it, the Hollow asks you: ``What name will you give in return?'' If you cannot give one, it takes a memory.

ℹtem[A] \textbf{The Unnamed Gate Key (Myth)} ℎfill \textit{[LEGEND] [ARTIFACT]}  
  A key made of brass and bone. It opens the Unnamed Gate on any night – not just solstice. The gate leads to the Hollow's domain. The key can be used once. After that, it turns to dust.  
  \textbf{Hollow cost:} Using the key invites the Hollow's attention. Tick \textbf{Unnumbered Grief} to maximum. The Hollow will follow you until you die – or until you give it a name it does not know.
\end{enumerate}

\section*{Quick use notes}
łabel{sec:galanina−quick−use}
\begin{itemize}
ℹtem Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = treasure. Highest rank sets the main Timer (2–5 $→$ 4, 6–10 $→$ 6, J/Q/K $→$ 8, A $→$ 10).
ℹtem Treasures (Diamonds) are codified outcomes (lamps, names, tears, keys) that change position rather than call for a roll.
ℹtem If any A appears, echo grief−omens: a green flame gutters, a child's shoe washes ashore, the ninth bell rings, the sea glows green at noon.
ℹtem Track the \textbf{Unnumbered Grief [8]} timer as a campaign−wide pressure. Use the \textbf{Hollow Walks} table when it reaches 8.
\end{itemize}

\subsection*{Additional Features}
\begin{itemize}
ℹtem \textbf{Lamp Oaths:} Swearing a promise over a green lamp binds the speaker. If broken, the lamp gutters and the Hollow visits the oath−breaker within nine days. The oath−breaker may appease the Hollow by lighting a new lamp for an unnumbered child (costs 1 Supply).
ℹtem \textbf{Green Flame Justice:} In Galanina, a witness may demand a trial by green flame. The accused holds a green lamp. If the flame burns true, they are innocent. If it gutters, they are guilty – and the Hollow takes them. No appeal.
ℹtem \textbf{Tears of the Vale:} Once per session, a character may collect tears from the olive grove (Wits+Survival DV 3). A tear can be used to calm the Hollow for one exchange or to restore 1 Fatigue (drinking it). The Hollow will remember the drinker.
ℹtem \textbf{The Unnumbered List:} The High Keeper maintains a list of names. A character who learns a name from the list may use it to bargain with the Hollow – the name can be offered in place of a memory or a life. The list has nine hundred names. The last nine are blank.
\end{itemize}

\subsection*{Lore Echoes (Cross−Expansion Hooks)}
\begin{itemize}
  ℹtem \textbf{The Brass Coast Sidhi Diaspora:} Galaninan green lamps are carried by Sidhi freedmen across the Amaranthine. A lamp from Galanina is a proof of ancestry; a lamp that gutters is a family's shame.
  ℹtem \textbf{The Covenant (Mykkiel):} The Covenant maintains a permanent garrison at Galanina. The judge there has authority to overrule any admiralty court in matters of slavery. The judge's silence is a vow – to speak would be to forget.
  ℹtem \textbf{Ashaan (The Veiled Throne):} Ashaan still claims Galanina as a lost province. Veiled Sisters send spies to retrieve the names of the children – to finish the ritual that would unseal the Hollow. The ritual requires nine names. The Sisters have eight.
  ℹtem \textbf{The Eastern Realms (The Hollow):} The Unnumbered Shore is considered a thin place by Aelaerem hedge−keepers. They leave green thread offerings there, hoping to guide the children home. The thread is always found wet – and knotted in a pattern no human tied.
  ℹtem \textbf{Saikou Ira's Compendium:} The Ninth Name Shrine is a known anchor for Oath−Keepers. Saikou Ira left a silver mirror there. The mirror shows not the viewer's face, but the face of the child they failed to protect. The mirror has nine cracks. The ninth crack weeps.
\end{itemize}

\subsection*{Galanina Superstitions (burn for +1 die once)}
\begin{itemize}
  ℹtem \textbf{Bread to the Flame:} Crumble a crust into a green lamp. Ignore the first \textit{Lamp Gutters} penalty.
  ℹtem \textbf{Knot the Wick:} Tie a single knot in a lamp's wick. Cancel \textit{Ashaani Bell Rings} or take –1 die on your next deception roll (your choice).
  ℹtem \textbf{Name to the Sea:} Whisper a dead child's name into the Unnumbered Shore. Gain +1 Effect when bargaining with the Hollow this scene, but mark \textit{Obligation} to the High Keeper.
\end{itemize}

\subsection*{Thematic SB Spend Table}
łabel{sec:galanina−sb}

⫕section*{Minor Complications (1 SB)}
\begin{itemize}
ℹtem \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Covenant marines or lamp−makers}.
ℹtem \textbf{Noise:} Sounds of your actions alert nearby \textbf{Hollow Children or lighthouse keepers}.
ℹtem \textbf{Trace:} Evidence of your passage marks your route for \textbf{the Oath−Keeper or Ashaani spies}.
ℹtem \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favourable tide}.
ℹtem \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

⫕section*{Moderate Setbacks (2 SB)}
\begin{itemize}
ℹtem \textbf{Alarm Raised:} \textbf{High Keeper or Covenant judge} becomes aware and begins responding.
ℹtem \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{fog or a green flash from the lighthouse}.
ℹtem \textbf{Foe Appears:} A \textbf{Hollow Child or Ashaani spy} arrives on scene.
ℹtem \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
ℹtem \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

⫕section*{Serious Trouble (3 SB)}
\begin{itemize}
ℹtem \textbf{Reinforcements:} Additional \textbf{Covenant marines, Hollow Children, or Ashaani agents} arrive.
ℹtem \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
ℹtem \textbf{Major Twist:} The situation fundamentally changes – \textbf{lamp gutters / sea weeps / ninth bell rings}.
ℹtem \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
ℹtem \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

⫕section*{Major Turns (4+ SB)}
\begin{itemize}
ℹtem \textbf{Trap Springs:} A prepared danger activates with full effect.
ℹtem \textbf{Authority Arrival:} \textbf{High Keeper Eliron, Covenant Judge Vered, or the Unnumbered Mother} intervenes.
ℹtem \textbf{Scene Shift:} The environment changes dramatically – \textbf{green flames turn blue / sea claims the shore / Unnamed Gate opens}.
ℹtem \textbf{Patron Omen:} Divine/arcane forces take notice – \textbf{omen appears / blessing lost / curse manifests}.
ℹtem \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

⫕section*{Region−Specific SB Options}
\begin{itemize}
ℹtem \textbf{Galanina (Green Flame):} Lamps flicker in unison, the sea reflects the lighthouse beam, a child's laughter echoes from the empty shore.
ℹtem \textbf{Galanina (Unnumbered):} You hear a name you thought you had forgotten; a child's footprint appears in your camp; the Hollow Children turn their heads toward you.
ℹtem \textbf{Galanina (Covenant Justice):} A clay tablet with your name appears on your bed; the judge's silent figure watches from a doorway; your shadow is missing a finger.
\end{itemize}

\subsection*{Boss Archetypes in Galanina}
łabel{sec:galanina−bosses}
\begin{itemize}
ℹtem \textbf{The Unnumbered Mother (Grief/Threshold)} – A grey figure who walks the shore calling names. \emph{Phases}: Calling $→$ Weeping $→$ Claiming. \emph{Cracks}: a \textbf{Green Lamp} lit for her lost child; a \textbf{Tear of the Vale} offered in silence; the name of the first child (found in the Sealed Archive).
ℹtem \textbf{The Hollow's Accountant (Debt/Justice)} – Faceless ledger−keeper who writes names. \emph{Phases}: Question $→$ Ledger Open $→$ Name Taken. \emph{Cracks}: a \textbf{Forgotten Name} given freely; a \textbf{Lamp Oath} sworn in its presence; a name from the Unnumbered List spoken without fear.
ℹtem \textbf{The Ashaani Bell−Ringer (Veiled Sister)} – Spy who seeks the ninth name. \emph{Phases}: Bell Tolls $→$ Children Stir $→$ Ritual Begins. \emph{Cracks}: expose her true face (she has none); destroy her brass bell; give her a false name – the Hollow will take her instead.
ℹtem \textbf{The Covenant Judge's Silence (Justice/Secrecy)} – A judge who has not spoken in nine years. \emph{Phases}: Tablet Written $→$ Witness Summoned $→$ Sentence Sealed. \emph{Cracks}: make her speak (requires a name from the Sealed Archive); burn her tablet; confess a crime she cannot judge (the Hollow will judge instead).
\end{itemize}

\subsection*{Faction Entanglements}
\begin{itemize}
  ℹtem \textbf{Covenant Garrison:} Respect lamp law (never extinguish a green flame) or face the judge's silence. Bringing a slaver to the Covenant reduces the next travel DV by 1.
  ℹtem \textbf{Lamp−Makers' Guild:} Commission a lamp (1 Supply) to gain +1 die on social rolls with Sidhi. Anger them (break a lamp oath) and the guild will refuse to sell you wicks – all lamp−based bonuses are lost.
  ℹtem \textbf{Hollow Children:} They do not speak, but they witness. A Hollow Child's pointing can be used as an omen (once per session). If you harm a Hollow Child, the Unnumbered Mother will call your name.
  ℹtem \textbf{Ashaani Remnant:} They trade in names. A name from the Sealed Archive can buy safe passage through Ashaani waters – but the Hollow will learn that name. The Veiled Sister's bell rings when she lies. It rings often.
  ℹtem \textbf{High Keeper's Circle:} Lighting a lamp for an unnumbered child (1 Supply) clears 1 Fatigue for all who witness. If you refuse to light a lamp when asked, the High Keeper will remember. The Hollow will too.
\end{itemize}

\subsection*{Cross−Regional Ties}
\begin{itemize}
  ℹtem \textbf{Brass Coast Sidhi Diaspora:} Galaninan green lamps are carried by Sidhi freedmen across the Amaranthine. A lamp from Galanina is a proof of ancestry; a lamp that gutters is a family's shame. The Sidhi of the Brass Coast will honour a Galaninan lamp as a letter of credit.
  ℹtem \textbf{Covenant (Mykkiel):} The Covenant garrison at Galanina reports directly to the High Synod. A Covenant Writ of Passage from Galanina is accepted in any Covenant court. The judge's silence is a vow – to speak would be to forget. The Covenant has forgotten many things.
  ℹtem \textbf{Ashaan (Veiled Throne):} The Veiled Sisters have tried to infiltrate Galanina nine times. The ninth attempt is ongoing. The spy's bell rings when she lies. It rings often. The Hollow listens.
  ℹtem \textbf{Eastern Realms (The Hollow):} The Unnumbered Shore is considered a thin place by Aelaerem hedge−keepers. They leave green thread offerings there, hoping to guide the children home. The thread is always found wet – and knotted in a pattern no human tied. The knot spells a name.
  ℹtem \textbf{Dhahara (Kalrantha Amulet):} The Sealed Archive contains a fragment of the Kalrantha amulet – a shard of green glass that the Ashaani used in their ritual. The shard is warm. It whispers the names of the children. The Covenant does not know it is there. The Hollow does.
\end{itemize}

\subsection*{Encounter Seeds (Quick Hooks)}
\begin{itemize}
ℹtem \textbf{The Ninth Lamp:} A lamp in the Green Bazaar begins to gutter. The lamp−maker refuses to touch it. The lamp's name is that of a PC's ancestor. The Hollow wants that name – or the PC's.
ℹtem \textbf{The Unnamed Gate Opens (Early):} The brass−and−bone door appears on the Unnumbered Shore. It is not solstice. A Hollow Child walks toward it. The party has nine rounds to stop the child – or to follow.
ℹtem \textbf{The High Keeper's Dream:} The High Keeper has a dream of a child's face. The child is alive – somewhere. She offers a \textbf{Green Lamp} to anyone who finds the child. The child is in Ashaan. The Veiled Sisters are watching.
ℹtem \textbf{The Oath−Keeper's Question:} The Oath−Keeper appears at the party's camp. It asks: ``What name did you forget?'' If the party cannot answer, it writes a name from the Unnumbered List – and that child will walk into the sea within the week.
\end{itemize}

% Index entries
∈dex{Galanina|see{Sealed Coast}}
∈dex{Sealed Coast generator}
∈dex{Places!Galanina}
∈dex{People!Galanina}
∈dex{Complications!Galanina}
∈dex{Rewards!Galanina}
∈dex{High Keeper Eliron}
∈dex{Lamp−Maker Shalvi}
∈dex{Covenant Judge Vered}
∈dex{Hollow Child}
∈dex{Unnumbered Shore}
∈dex{Ninth Name Shrine}
∈dex{Green lamp}
∈dex{Unnumbered Grief}
∈dex{Lamp etiquette}
∈dex{Hollow Walks}
