\section{Sihai --- ``The Central Kingdom''}
\label{chap:sihai}

\textbf{Key Demographics: } Humans
\textbf{Capital City: } Jingdu

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The Mandate is a memory. The ancestors are listening. The Hollow collects what is forgotten.}

\begin{quote}
  \textbf{Quiet Chancellor (Elite):}  
  ``The Mandate of Heaven is not a right. It is a story we tell ourselves --- that the emperor remembers the names of the dead. When the story falters, the empire falls. I do not serve the emperor. I serve the silence where the forgotten names should be.''
\end{quote}

\begin{quote}
  \textbf{Provincial Censor (Commoner):}  
  ``My brush is heavier than any sword. A single stroke can make a name vanish --- from ledgers, from lips, from the dreams of those who loved it. The emperor reads my reports and trembles. He knows that I have written nine reports this year. The ninth is about him.''
\end{quote}

\noindent\textbf{Genre / Mood:} Celestial bureaucracy, ink-washed horror, examination hell, the weight of forgotten ancestors.

\noindent\textbf{Starting Location:} An examination hall where ink-stained desks face a silent proctor. A Provincial Censor holds up a scroll: ``This candidate's family name has been erased. He does not know. You will tell him --- or you will take his place in the silence.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
Sihai is the central kingdom, the seat of the Emperor and the Mandate of Heaven. Its canals carry grain from the south; its examination halls produce bureaucrats who govern by the brush. But beneath the mortal bureaucracy runs a deeper one: the Celestial Bureaucracy of gods, ghosts, and ancestors, where every official has a heavenly counterpart and every erased name becomes a hungry ghost. The Lethai-thora --- the ancient archivists who taught the first emperors to write --- still keep the hidden scrolls, the ones that record what the Censorate burns.

Power here is measured in ink, not iron. The Censorate watches everyone; the Salt Monopoly controls trade; the Kitchen God reports every household's secrets to the Jade Emperor at New Year. A well-written couplet can appease a ghost; a forged seal can summon a demon. The Hollow's ninth weight is the silence at the heart of the Mandate itself. Every dynasty inherits its legitimacy; every silence is a weight.

\vspace{2mm}
\textbf{What you notice first:} The smell of ink and canal water, but also incense and grave-soil. The way officials stamp every document with a brass seal --- and the way the seal leaves a faint watermark that looks like a face. The constant ticking of water-timers, but sometimes the ticking skips a beat, and the shadows pause to listen.

\textbf{The one rule that kills newcomers:} Never speak of succession before naming the Emperor. The Censorate has ears in every tea house --- and the ghosts have ears in every wall. A misplaced word can make a name vanish from the rolls, and from the memory of the ancestors.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Brush}

Power here is measured in examination scores, salt licenses, and the thickness of a family's archive. The Emperor rules, but the Censorate audits, the Salt Monopoly finances, and the Lethai-thora remember what the Censorate burns. A provincial governor's authority rests on his examination rank; a merchant's on his salt license. The dead are not gone --- they are merely demoted. The Hungry Ghost Festival is the only time the Celestial Bureaucracy accepts appeals. The Hollow watches from the missing ninth step, the unpoured ninth cup, the unsaid ninth name.

\subsection*{Prominent NPCs of Sihai}
\label{sec:sihai-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Quiet Chancellor Wei & Imperial censor & Root out corruption before the Mandate fails & He has never smiled. His seal is a laughing mouth. He speaks to the Kitchen God each evening. \\
\hline
Salt Commissioner Li & Head of the Salt Monopoly & Control the price of salt and the flow of silver & She tastes every batch of salt before it is stamped. Her tongue is white. She sees ghosts in the brine. \\
\hline
Provincial Censor Chen & Overseer of the eastern prefectures & Erase a rebel lineage from the rolls & He writes with his left hand. His right hand is for swearing oaths. His left hand has no fingerprints. \\
\hline
Lethai-thora Keeper Lian & Hidden archivist & Protect a scroll that proves the emperor's claim is false & She has not spoken aloud in twenty years. She writes on water. The water never freezes. \\
\hline
The Hollow's Oath-Keeper & Spirit of the fractured Mandate & Collect the ninth name, the ninth step, the ninth cup & It appears as a scribe with too many fingers. It never blinks. It smells of old incense. \\
\hline
\end{longtable}

\subsection*{Names of Sihai}
\label{sec:sihai-names}

Inspired by the central kingdom: short, tonal, with honorifics. Surnames often denote a clan or a prefecture.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/House} & \textbf{Epithets} \\
\hline
Wei & Lian & Chen & \textit{the Quiet} \\
\hline
Li & Mei & Zhou & \textit{Salt-Tongue} \\
\hline
Chen & Huan & Zhao & \textit{the Unerased} \\
\hline
Zhang & Ning & Wu & \textit{Ninth Seal} \\
\hline
Liu & Ping & Tang & \textit{the Brush} \\
\hline
\end{longtable}

\paragraph*{(Hall/Canal/Archive)} Examination hall with ink-stained desks; Grand Canal lock with barges queueing; Lethai-thora hidden archive behind a waterfall.

\section*{Spades --- Places}
\label{sec:sihai-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Examination Hall (Hell of Ink)} --- A long building with ink-stained desks, silent proctors, and a single bell.
  \hfill \textit{The bell rings when a candidate cheats. It rings once a day on average. The ghosts of failed candidates write answers on the ceiling.} \texttt{[DEATH]}
\item \textbf{Grand Canal Lock (Dragon Gate)} --- A stone gate where barges queue for inspection. Officials stamp manifests.
  \hfill \textit{The stamp leaves a watermark that glows under moonlight --- a dragon scale. Forgeries are obvious. The dragon knows.} \texttt{[MAGIC]}
\item \textbf{Lethai-thora Hidden Archive} --- A pavilion behind a waterfall, filled with forbidden scrolls.
  \hfill \textit{The scrolls are sealed with silk cords. The cords have nine knots. The ninth is a lie. The archivists are all ghosts.} \texttt{[SECRET]}
\item \textbf{Salt Commissioner's Counting House} --- A fortress of ledgers, guarded by armed clerks.
  \hfill \textit{The ledgers are written in disappearing ink. Only the Commissioner knows the reagent --- tears of a famine widow.} \texttt{[AUTHORITY]}
\item \textbf{Provincial Censorate Office} --- A gray building with a brass scale above the door.
  \hfill \textit{The scale weighs truth and lies. It has never balanced. When it tips, someone is erased.} \texttt{[JUSTICE]}
\item \textbf{Emperor's Forbidden Garden (Kunlun Pavilion)} --- A walled garden where no one may enter without permission.
  \hfill \textit{The garden contains a single tree with nine branches. The ninth branch is dead. A phoenix nests in the eighth.} \texttt{[MYTH]}
\item \textbf{Well of Erased Names} --- A stone well where the Censorate throws scrolls of erased lineages.
  \hfill \textit{The scrolls burn as they fall. The smoke smells of ink and old blood. The well is bottomless.} \texttt{[DEATH]}
\item \textbf{Tea House of the Ninth Cup (Ghost Festival)} --- A popular establishment where the ninth cup is never poured.
  \hfill \textit{The owner has a bell that rings when someone pours the ninth cup. It rings often. The ring attracts hungry ghosts.} \texttt{[SOCIAL]}
\item \textbf{Canal Mother's Court} --- A shaded bench where water rights are assigned.
  \hfill \textit{The bench is carved with the names of every canal mother for nine generations. The ninth name is blank. It waits.} \texttt{[COMMUNITY]}
\item[\card{\spadesuit}{J}] \textbf{The Emperor's Tomb (Unfinished)} --- A subterranean complex where the current emperor's tomb is being built.
  \hfill \textit{The workers are slaves. They will be buried with him. They know. They carve their names into the walls.} \texttt{[DEATH]}
\item[\card{\spadesuit}{Q}] \textbf{The Hollow's Gate (Subterranean)} --- A crack in the floor of the Forbidden Garden.
  \hfill \textit{The crack is warm. It whispers the names of emperors who died without heirs. The names are getting longer.} \texttt{[MYSTERY]}
\item[\card{\spadesuit}{K}] \textbf{The Ninth Ledger's Vault} --- A vault beneath the Counting House, accessible only by nine keys.
  \hfill \textit{The ninth key is held by the Hollow. The Hollow will surrender it --- for a price. The price is a name you have never spoken.} \texttt{[WEIGHT]}
\item[\card{\spadesuit}{A}] \textbf{The First Examination Hall (Myth)} --- Where the first emperor tested his sons.
  \hfill \textit{The hall is invisible. Only those who have passed nine examinations can see it. Inside, the first emperor still grades papers.} \texttt{[LEGEND]}
\end{enumerate}

\paragraph*{(Censor/Merchant/Keeper)} Provincial censor with a list of forbidden words; salt merchant with honest scales except when they aren't; Lethai-thora memory-keeper with a silence-token.

\section*{Hearts --- People \& Factions}
\label{sec:sihai-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Provincial Censor} --- Gray-robed, a brass seal on a chain. He carries a list of forbidden words.
  \hfill \textit{``Speak the ninth word,'' he says, ``and your name will join the list.'' The list is written on human skin.} \texttt{[AUTHORITY]}
\item \textbf{Salt Merchant} --- A thin woman with a brass scale. Her weights are hollow.
  \hfill \textit{``The scale is honest,'' she says. ``My hand is not.'' She has nine weights. The ninth is empty.} \texttt{[TRADE]}
\item \textbf{Lethai-thora Keeper} --- Silent, carrying a brass lamp. She writes on water with a reed brush.
  \hfill \textit{``The words vanish,'' she says. ``That is how I know they are true.'' Her lamp burns without oil.} \texttt{[SECRET]}
\item \textbf{Canal Mother} --- A matriarch in blue robes, a key to the sluice gate on her belt.
  \hfill \textit{``Water is memory,'' she says. ``The canal remembers every bribe.'' She has never taken a bribe. She has taken nine stories.} \texttt{[COMMUNITY]}
\item \textbf{Examination Candidate} --- A young man with ink-stained fingers. He has failed nine times.
  \hfill \textit{``The tenth will be different,'' he says. He has said this nine times. His shadow is missing two fingers.} \texttt{[DESPERATION]}
\item \textbf{Quiet Chancellor's Spy} --- A woman disguised as a tea-seller. She listens at every table.
  \hfill \textit{She writes reports on her own skin. The ink is invisible. Her skin is full. She is running out of space.} \texttt{[STEALTH]}
\item \textbf{Salt Worker (Slave)} --- A man with cracked hands, harvesting salt from pans.
  \hfill \textit{``The salt tastes of tears,'' he says. ``Whose tears, I do not know.'' He is bound to nine years of labor. He has served eight.} \texttt{[LABOR]}
\item \textbf{Hollow's Oath-Keeper} --- A tall figure in gray, with too many fingers. It carries a ledger.
  \hfill \textit{It asks: ``Where is the ninth name?'' It will not leave without an answer. Its ledger is made of human bone.} \texttt{[DEATH]}
\item \textbf{Emperor's Dreamer} --- A woman locked in a tower. She dreams the emperor's nightmares.
  \hfill \textit{She wakes screaming. The dreams are always about the Hollow. She has not dreamed in nine days. The emperor has not slept.} \texttt{[MYSTERY]}
\item[\card{\heartsuit}{J}] \textbf{The First Censor (Ghost)} --- A skeleton in gray robes, sitting in the Well of Erased Names.
  \hfill \textit{He asks: ``Did I erase the right name?'' The answer is always a lie. He knows. He asks anyway.} \texttt{[DEATH]}
\item[\card{\heartsuit}{Q}] \textbf{Canal Mother's Ghost} --- A translucent figure walking the canal at midnight.
  \hfill \textit{She checks the water level. When it drops, a reckoning is due. She carries a dipper that never empties.} \texttt{[COMMUNITY]}
\item[\card{\heartsuit}{K}] \textbf{The Ninth Emperor's Ghost} --- A crowned skeleton in the unfinished tomb.
  \hfill \textit{He asks: ``Is the Mandate renewed?'' If you say no, he weeps. His tears turn to ink. The ink writes new laws.} \texttt{[LEGEND]}
\item[\card{\heartsuit}{A}] \textbf{Kuva's Shadow (Hollow)} --- The vast darkness that covers the Forbidden Garden at midnight.
  \hfill \textit{It asks: ``Who holds the ninth key?'' The answer is a name no one knows. The darkness absorbs the answer.} \texttt{[PATRON]}
\end{enumerate}

\paragraph*{(Forgery/Monsoon/Ancestor)} A scroll is found to be a forgery --- the party is blamed; monsoon breaks early --- the canal closes; an ancestor demands a public accounting of a forgotten burden.

\section*{Clubs --- Complications/Threats}
\label{sec:sihai-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Forged Scroll} --- A tax ledger is discovered to be a fake. The forgery is perfect.
  \hfill \textit{The forger is a Censorate clerk. The scroll is not a forgery --- it is the original. The Censorate has been lying for centuries.} \texttt{[DECEPTION]}
\item \textbf{Monsoon Breaks Early} --- The canal closes for inspection. Delay 2 segments.
  \hfill \textit{The closure is a cover. The Censorate is searching for a specific barge. The barge carries a forbidden ancestor tablet.} \texttt{[WEATHER]}
\item \textbf{Ancestor's Demand} --- A ghost appears at the family shrine. It demands a public accounting.
  \hfill \textit{The obligation is from nine generations ago. The ancestor will not leave until it is settled. It will accept a story if the story is true.} \texttt{[DEATH]}
\item \textbf{Censorate Audit} --- Officials arrive with a warrant. They have a list of forbidden words.
  \hfill \textit{The list includes a word you spoke yesterday. The penalty is erasure. The word was your grandmother's name.} \texttt{[AUTHORITY]}
\item \textbf{Salt Monopoly Price Hike} --- The Commissioner doubles the salt tax. Riots begin.
  \hfill \textit{The hike is a bribe. The bribe is to the Quiet Chancellor. He accepted. The salt is now gray. It used to be white.} \texttt{[TRADE]}
\item \textbf{Canal Lock Sabotage} --- Someone drops a barge at the lock. The canal backs up.
  \hfill \textit{The saboteur is a Canal Mother's son. He wants to flood the Censorate office. The water is rising. The papers are drowning.} \texttt{[DISASTER]}
\item \textbf{Examination Cheating Scandal} --- A candidate is caught with a cheat sheet. The sheet has your name.
  \hfill \textit{The cheat sheet was planted. The real target is your patron. The cheat sheet is written in disappearing ink. It will be gone by dawn.} \texttt{[TRAP]}
\item \textbf{Lethai-thora Archive Fire} --- A lamp falls in the hidden pavilion. Scrolls burn.
  \hfill \textit{The fire was set by the Censorate. They want to erase a specific history. The history is about the first emperor's true lineage.} \texttt{[DESTRUCTION]}
\item \textbf{Hollow's Weight Called} --- An Oath-Keeper appears at the tea house. It points at a party member.
  \hfill \textit{A burden has found its bearer. The burden is the ninth cup. Settle it in story or memory. The Oath-Keeper will know if you lie.} \texttt{[WEIGHT]}
\item[\card{\clubsuit}{J}] \textbf{Emperor's Health Fails} --- The Mandate wavers. Succession rumors spike.
  \hfill \textit{The emperor has no heir. The Hollow is watching. It will claim the ninth name. The name is the emperor's. He does not know it.} \texttt{[POLITICS]}
\item[\card{\clubsuit}{Q}] \textbf{Forbidden Garden Breached} --- Someone enters the garden without permission. The Hollow follows.
  \hfill \textit{The intruder is a child. The Hollow asks: ``Why are you here?'' The child does not know. The child is a ghost. It has been dead for nine years.} \texttt{[MYSTERY]}
\item[\card{\clubsuit}{K}] \textbf{Ninth Ledger Stolen} --- The vault is cracked. The ninth ledger is missing.
  \hfill \textit{The ledger contains the names of everyone the Censorate has erased. The thief is a ghost. The ghost is the first emperor. He wants his name back.} \texttt{[DEATH]}
\item[\card{\clubsuit}{A}] \textbf{Mandate Falls} --- The emperor dies without an heir. The Hollow claims the throne.
  \hfill \textit{The Hollow sits on the dragon throne. It asks: ``Who will rule?'' Silence is the only answer. The Hollow begins to write.} \texttt{[APOCALYPSE]}
\end{enumerate}

\paragraph*{(Pass/Seal/Token)} Stamped pass from a provincial governor (one use); copy of the Celestial Ledger (reveals one hidden tax); Lethai-thora silence-token (one question the keeper must answer).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:sihai-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Stamped Pass (Governor's Seal)} --- A wax-sealed document. Grants passage through any provincial checkpoint.
  \hfill \textit{The seal is a tiger. It glows when a Censorate agent is nearby. The tiger's eyes follow you.} \texttt{[TRAVEL]}
\item \textbf{Celestial Ledger (Copy)} --- A scroll containing one hidden tax. Use it to blackmail an official.
  \hfill \textit{The ledger is written in disappearing ink. Read it aloud before it fades. The ink tastes of copper.} \texttt{[BLACKMAIL]}
\item \textbf{Lethai-thora Silence-Token} --- A brass charm. Present it to demand a keeper answer one question.
  \hfill \textit{The keeper cannot lie. The question must be about a forgotten secret. The secret is never small.} \texttt{[SOCIAL]}
\item \textbf{Junk Seal (Official)} --- A brass seal. Impersonate a provincial official for one scene.
  \hfill \textit{The seal is warm. It glows when someone tests its authenticity. The glow is the color of a lie.} \texttt{[DECEPTION]}
\item \textbf{Canal Mother's Key} --- A bronze key. Open any sluice gate in the Grand Canal.
  \hfill \textit{The key is cold. It warms when a water right is being violated. The warmth is the temperature of a drowning.} \texttt{[AUTHORITY]}
\item \textbf{Salt Commissioner's Waiver} --- A stamped document. Exempts you from the salt tax for one year.
  \hfill \textit{The waiver is written on silk. The silk was stolen from the Commissioner's own robes. The thread is from a noose.} \texttt{[TRADE]}
\item \textbf{Examination Answer Key (Forged)} --- A cheat sheet that actually works. Pass one examination automatically.
  \hfill \textit{The answers are written in invisible ink. The reagent is moonlight. The answers are wrong --- but the examiners cannot tell.} \texttt{[DECEPTION]}
\item \textbf{Quiet Chancellor's Seal (Jade)} --- A jade seal. Commands the Censorate to ignore one crime.
  \hfill \textit{The seal is a forgery. The Chancellor made it himself. It still works. The jade is warm. It was cut from a tomb.} \texttt{[AUTHORITY]}
\item \textbf{Forbidden Garden Key (Iron)} --- A heavy iron key. Opens the gate to the Emperor's garden.
  \hfill \textit{The key is cursed. Each use ages the wielder one month. The key does not care. It has been waiting for centuries.} \texttt{[MYSTERY]}
\item[\card{\diamondsuit}{J}] \textbf{Hollow's Ninth Ledger Page} --- A single page from the ninth ledger. Contains nine erased names.
  \hfill \textit{Speak a name, and that person's deeds become known. The Hollow will remember too. The page is warm. It has a pulse.} \texttt{[DEATH]}
\item[\card{\diamondsuit}{Q}] \textbf{First Censor's Brush} --- A brush made from a phoenix feather. Write any name, and it is erased from all records.
  \hfill \textit{The brush works once. Then it crumbles. The Hollow will know. The phoenix will not forgive.} \texttt{[LEGEND]}
\item[\card{\diamondsuit}{K}] \textbf{Canal Mother's Ghostly Dipper} --- A clay dipper used by a dead Canal Mother. Use it to control any sluice gate.
  \hfill \textit{The dipper is cold. It weeps water when a reckoning is due. The water is salt. It tastes of old grief.} \texttt{[COMMUNITY]}
\item[\card{\diamondsuit}{A}] \textbf{The Mandate Itself (Scroll)} --- A scroll that proves the current emperor has no legitimate claim.
  \hfill \textit{Read it aloud, and the Mandate falls. The Hollow will take the throne. The scroll burns after reading. The smoke is a new name.} \texttt{[APOCALYPSE]}
\end{enumerate}

% ============================================================
% ADDITIONAL GENERATORS (from The Eastern Realms expansion)
% ============================================================

\subsection*{Censorate Audit Complications (d6)}
\label{sec:sihai-audit-complications}

When a character faces a Censorate audit (see Additional Features), roll or choose an additional complication that escalates the stakes.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d6} & \textbf{Complication} \\
\hline
1 & \textbf{Erased Name} --- The Censorate has struck your name from a ledger. You are legally invisible. Gain $-1$ Position on all social rolls with officials until resolved. \\
\hline
2 & \textbf{Ancestor Summons} --- A ghost appears at the family shrine demanding a public accounting. The burden is from nine generations ago. \\
\hline
3 & \textbf{Forged Scroll} --- A tax ledger is discovered to be a forgery. The party is blamed. The real forger is a Censorate clerk. \\
\hline
4 & \textbf{Quiet Chancellor's Interest} --- The Chancellor requests a private audience. Refusal is treason. Attendance incurs a favor. \\
\hline
5 & \textbf{Lethai-thora Silence-Token} --- A hidden archivist offers a token in exchange for a secret. The secret will be witnessed by the dead. \\
\hline
6 & \textbf{Well of Erased Names} --- A character sees their own name floating in the well. The Hollow is watching. Gain the \textit{Marked} condition. \\
\hline
\end{longtable}

\subsection*{Hidden Archive Discoveries (d8)}
\label{sec:sihai-archive-discoveries}

When the party explores the Lethai-thora Hidden Archive or the Ninth Ledger's Vault, roll to determine what forgotten secret they uncover.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d8} & \textbf{Discovery} \\
\hline
1 & A scroll that proves the current emperor's claim is false. \\
\hline
2 & A map of the Ninth Ledger's vault, sealed with a finger bone key. \\
\hline
3 & A forgotten ancestor tablet still warm to the touch. \\
\hline
4 & The true name of a Veiled Sister, written in disappearing ink. \\
\hline
5 & A Lethai-thora memory-thread containing the last breath of a dead emperor. \\
\hline
6 & The original Mandate of Heaven, showing the ninth clause was added later. \\
\hline
7 & A canal mother's key that opens any sluice gate --- and a ghost that wants it back. \\
\hline
8 & A blank scroll that writes itself when no one is watching. It is writing your obituary. \\
\hline
\end{longtable}

\subsection*{Water Rights Disputes (d6)}
\label{sec:sihai-water-disputes}

When a conflict arises over canal access or irrigation, roll to determine the nature of the dispute.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d6} & \textbf{Dispute} \\
\hline
1 & A Canal Mother accuses a rival of diverting water after dark. The evidence is a single wet footprint. \\
\hline
2 & The Dragon King is offended. A barge carrying a forbidden ancestor tablet has contaminated the lock. \\
\hline
3 & A salt merchant claims the right to dam a tributary for his pans. The farmers downstream are sharpening sickles. \\
\hline
4 & The water level dropped overnight. A ghostly Canal Mother walks the lock, demanding to know who stole her key. \\
\hline
5 & Two examination candidates accuse each other of bribing the lockmaster. The lockmaster has vanished. \\
\hline
6 & The Ninth Lock has frozen in summer. The Hollow is breathing through the ice. No one will go near it. \\
\hline
\end{longtable}

\section*{Quick use notes}
\label{sec:sihai-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer ($2$--$5$ $\rightarrow$ $4$, $6$--$10$ $\rightarrow$ $6$, J/Q/K $\rightarrow$ $8$, A $\rightarrow$ $10$).
\item Diamonds are codified outcomes (passes, seals, tokens) that change position rather than call for a roll.
\item If any A appears, echo \textbf{ink-omens}---scrolls that rustle in still air, a lamp that burns green, a forgotten name whispered from a shelf, a ghost that asks for a pen.
\end{itemize}

\subsection*{Additional Features (Celestial Bureaucracy)}
\label{sec:sihai-features}

\begin{itemize}
\item \textbf{The Mandate Timer [8] (Shared Campaign Timer ):} Tracks the stability of the dynasty. The party inherits 2 ticks at the start of any adventure set in Sihai (the inherited burden). The timer ticks when the emperor's health fails, when a succession dispute arises, when a major corruption scandal is exposed, or when the party breaks an oath witnessed by an ancestor. When the timer reaches 4: the Censorate becomes more aggressive (all audits start at Desperate Position). When it reaches 8: the Mandate fails --- the Hollow claims the throne, civil war erupts, and the party must choose a side or flee. Resetting the timer requires an \textbf{Ancestor Rite} at the Imperial Ancestral Temple (downtime, costs 1 Supply and a truthful confession) or a public execution of a corrupt official (costs 1 Boon and may create enemies).

\item \textbf{The Two Ledgers (Ancestral Names Mechanic):} Every character has two ``names'': their \textbf{Mortal Name} (what the Censorate knows) and their \textbf{Ancestral Name} (recorded in the Celestial Ledger, kept by the Lethai-thora). The Ancestral Name cannot be erased by mortal means. Discovering a character's Ancestral Name requires a \textbf{Research Timer (size 6)} in the Lethai-thora archives. Once known, the player may invoke the Ancestral Name once per session to gain $+2$ dice on any roll related to that ancestor's domain (e.g., a farmer's ancestor aids harvest, a soldier's ancestor aids battle). However, invoking the Ancestral Name creates an \textbf{Ancestral Weight (size 4 timer)}. When it fills, the ancestor demands a service --- usually something the ancestor failed to do in life (e.g., ``Apologize to my widow'' or ``Find my stolen seal''). Refusing the demand marks 1 Corruption and adds 2 ticks to the Mandate Timer .

\item \textbf{Censorate Audit (Social Combat Mechanic):} The Censorate can audit any character at any time. An audit is a social combat with three phases:
  \begin{enumerate}
    \item \textbf{Pleadings} (Presence + Sway vs.\ Censor's Lore). Success grants $+1$ die to the next phase.
    \item \textbf{Evidence} (Wits + Investigation vs.\ DV 4). Success reveals a hidden burden or a forged document; failure gives the Censor a piece of evidence against you.
    \item \textbf{Judgment} (Resolve + Lore vs.\ DV 5). Success: the character is cleared. Failure: the Censorate erases a name --- the character loses one Bond (that person no longer remembers them) or one Asset (it is seized), and the Mandate Timer ticks $+1$.
  \end{enumerate}
  Players can avoid an audit by: \textbf{Bribing the Censor} (costs 1 Honey Jar or 1 Boon), \textbf{Presenting a forged document} (requires a Forgery Timer , size 4), or \textbf{Calling a ghost witness} (requires a Spirit + Lore test, DV 4; the ghost cannot lie, but it will ask for a favor in return -- usually a promise that creates a new Ancestral Weight).

\item \textbf{Examination Hell (Dungeon Mechanic):} The Forbidden Garden contains a subterranean examination hall where the ghosts of failed candidates still write their answers. The party can enter to find a lost Ancestral Name -- but the hall is alive.
  \begin{itemize}
    \item The hall has a \textbf{Stress Timer (size 8)}. Each failed roll, each lie spoken, each broken taboo ticks the timer.
    \item At 4: The hall begins to change. Doors lead to different corridors. Navigation becomes Desperate.
    \item At 8: The \textbf{Examiner Ghost} appears --- a skeleton in scholar's robes. It asks one question. The party must answer in verse. If the verse is good, the ghost grants a boon (reveals one Ancestral Name). If poor, the ghost erases one of the party's names --- that character is forgotten by the living for $1\text{d}4$ days (lose all Bonds temporarily, cannot be recognized). The Stress Timer resets after the encounter.
  \end{itemize}

\item \textbf{Paper Soldiers and Hungry Ghosts (Minion Types):}
  \begin{itemize}
    \item \textbf{Paper Soldiers (Tier I, Cap 2):} Origami constructs animated by Censorate magic. They have Harm 2, Body 2, Melee 1. They are Vulnerable to fire (double Harm) and to a successful Lore test (reveals their true name, causing them to crumple).
    \item \textbf{Hungry Ghosts (Tier II, Cap 3):} Spirits of those whose names were erased. They have Harm 3, Spirit 3, and the ability \textbf{Memory Drain} (touch, target forgets a minor detail --- a face, a password, a skill point for one scene). They cannot cross a line of salt or a properly witnessed oath-circle.
  \end{itemize}

\item \textbf{Water Rights as Ancestral Currency:} In the canal system, water rights are traded like silver, but they are also offerings to the Dragon Kings of the Four Seas. A Canal Mother's key is worth more than a chest of coins --- it can also summon a rain spirit (once per season, calm a storm or flood a district). Forging a water right angers the Dragon King; the forger's wells run dry for nine years (all Survival tests to find water suffer $-2$ dice).

\item \textbf{Lethai-thora Archivists of the First Seal:} The Lethai-thora are not merely librarians --- they are the keepers of the original Mandate, written on tortoise shells and dragon bones. They do not serve the emperor; they serve the record that predates the empire. A Lethai-thora keeper can declare any document void by writing ``void'' in water over it. The water must be from the Well of Erased Names. Doing so costs 1 Obligation to the Lethai-thora and ticks the Mandate Timer $+1$ (the Censorate will investigate).

\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Jade Censor} --- The Witness. \textit{A magistrate who never blinks, who writes every truth in a ledger of jade.}
    \item \textbf{The Dragon Salt} --- Maelstraeus. \textit{The price of a monopoly; the hidden cost of a nation's thirst.}
    \item \textbf{The Ninth Ancestor} --- Khemesh. \textit{The ghost of the first emperor, who still reviews petitions.}
    \item \textbf{The Brush That Walks} --- The Traveler. \textit{A calligraphy brush that guides lost officials through the underworld.}
    \item \textbf{The Kitchen's Ear} --- The Sealed Gate. \textit{The god who reports every secret to the Celestial Bureaucracy.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Ninth Ledger and the Kalrantha Amulet (Dhahara):} The ninth ledger is said to contain the name of the prince who stole the amulet. The Veiled Sisters want to erase it. The Breath-less want to find it. The Lethai-thora have hidden it in a scroll that unrolls forever.
  \item \textbf{Silence-Tokens and the Covenant (Mykkiel):} The Covenant recognizes Lethai-thora silence-tokens as valid requests for closed hearings. A silence token can pause any court proceeding for the duration of a single breath. The token is a strip of silk from a burial shroud.
  \item \textbf{Canal Mothers and the Fhara (Way of Silk):} Fhara well-judges and Sihai canal mothers use similar water-measuring systems. A canal mother's key can be used to negotiate a well contract --- but the key must be dipped in the well first. The water will remember.
  \item \textbf{The Mandate and the Hollow (Eastern Realms):} The Hollow is the silence at the heart of the Mandate. Aelaerem hedge-keepers say that when a dynasty falls, the Hollow eats the ninth name. The name is always the emperor's secret childhood name.
\end{itemize}

\subsection*{Sihai Superstitions (burn for $+1$ die once)}
\begin{itemize}
  \item \textbf{Bread to the Canal:} Crumble a crust into the water. Ignore the first \emph{Forged Scroll} penalty. \texttt{[SUPERSTITION]}
  \item \textbf{Knot the Scroll:} Tie a single knot in a silk cord. Cancel \emph{Censorate Audit} or take $-1$ die on your next forgery detection roll (your choice). \texttt{[SUPERSTITION]}
  \item \textbf{Name to the Well:} Whisper a dead censor's name into the Well of Erased Names. Gain $+1$ Effect when negotiating with the Censorate this scene, but mark \emph{Obligation} to the Hollow. \texttt{[SUPERSTITION]}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:sihai-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Censorate clerks or ancestor ghosts}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{canal patrols or hungry ghosts}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{trackers or the Kitchen God}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable water level}.
\item \textbf{Supply Strain:} Mark $+1$ segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local censor or canal mother} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{fog or a ghost procession}.
\item \textbf{Foe Appears:} A \textbf{paper soldier (origami construct) or a hungry ghost} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Censorate guards, ancestor spirits, or paper soldiers} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes---\textbf{ancestor summoned/ledger stolen/mandate questioned}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns ($4+$ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Quiet Chancellor, Salt Commissioner, or First Censor's Ghost} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically---\textbf{canal floods/archive burns/mandate falls}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice---\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Sihai (Celestial Bureaucracy):} A ghost clerk appears with a writ, a paper seal floats down from nowhere, the Emperor's Dreamer whispers a warning.
\item \textbf{Sihai (Ancestor Law):} A family shrine glows, an ancestor's portrait weeps ink, a forgotten burden is recalled in a dream.
\item \textbf{Sihai (Examination Hell):} A failed candidate's ghost asks for a pen, a test paper bleeds, the proctor's bell rings nine times.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Navigation):} On a hit, shift one band. On a strong hit, also impose \emph{Paper Trail}.
\item \textbf{Cut the Seal (Presence + Lore):} On a hit, freeze range for one exchange; if a stamped pass is in your possession, $+1$ die.
\item \textbf{Weather the Audit (Resolve + Spirit):} On a hit in \emph{Censorate Audit}, you pick who testifies; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Brush \& Mandate)}
\begin{itemize}
\item \textbf{The Forgotten Emperor (Ghost of the Ninth Seal)} --- \emph{Phases}: Names Erased $\rightarrow$ Ledger Stolen $\rightarrow$ Mandate Reclaimed. \emph{Cracks}: present a true ancestor tablet, expose a forged lineage, survive a night in the Well of Erased Names.
\item \textbf{The Brush That Writes Itself (Living Ink)} --- \emph{Phases}: Paper Soldiers $\rightarrow$ Scroll Prison $\rightarrow$ Ink Flood. \emph{Cracks}: a Lethai-thora silence-token, a confession written in water, the First Censor's Brush.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power},breakable]
In Sihai, power flows through ink, salt, and the weight of the Mandate. The Emperor rules, but the Censorate audits, the Salt Monopoly finances, the Lethai-thora remember, and the ancestors watch. A well-written couplet can appease a ghost; a forged seal can start a war. The Celestial Bureaucracy is always hiring.

\textbf{For the GM:}  
Power in Sihai is measured in examination scores, salt licenses, and the thickness of a family's archive. Patronage often takes the form of stamped passes, ledger copies, or silence-tokens --- each tied to a specific office, canal, or burden. To emphasize the mythological flavor:
\begin{itemize}
\item Use the \textbf{Mandate Timer } as a visible campaign front. The party's actions can tick or reset it. Ancestor rites are the only safe way to reduce it.
\item Enforce \textbf{Censorate Audits} as a recurring threat --- players must maintain clean papers or face erasure. Ghost witnesses can be summoned by either side.
\item Let \textbf{water rights} be a form of currency that can be stolen, forged, or contested --- but angering the Dragon King has consequences.
\item Introduce \textbf{paper soldiers} (origami constructs) as low-tier enemies and \textbf{hungry ghosts} as mid-tier threats. The Censorate employs both.
\end{itemize}
In Sihai, the brush forgets nothing --- and the Hollow remembers every erased name.
\end{tcolorbox}

% Index entries
\index{Sihai|see{Central Kingdom}}
\index{Theme generator}
\index{Sihai generator}
\index{Places!Sihai}
\index{People!Sihai}
\index{Complications!Sihai}
\index{Rewards!Sihai}
\index{Celestial Bureaucracy}
\index{Censorate}
\index{Salt Monopoly}
\index{Lethai-thora}
\index{Mandate of Heaven}
\index{Hollow's Oath-Keeper}
\index{Dragon King}
\index{Kitchen God}
\index{Paper soldiers}
\index{Hungry ghosts}
\index{Examination hell}
\index{Well of Erased Names}
\index{Ninth Ledger}
\index{Local Patrons!Sihai}