\section{Taharka — ``The Monsoon Crown''}
\label{chap:taharka}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The mountain listens. The horse remembers. The monsoon keeps the balance.}

\begin{quote}
  \textbf{Ridge-Watcher Captain (Elite):}  
  ``The passes remember every army that tried to climb them. The mountains do not forgive haste, and they do not forget a foreign banner. We do not guard the highlands. We listen to them. They tell us who to let through.''
\end{quote}

\begin{quote}
  \textbf{Horse Trader (Commoner):}  
  ``Before you praise the rider, praise the horse. Before you praise the horse, praise the mountain that grew its lungs. A Taharkan foal is born breathing thin air, and each foal carries a \textit{fumbo} – a riddle from its ancestor's death. To ride it, you must answer. To answer wrong is to fall. To answer true is to be remembered.''
\end{quote}

\textbf{Genre / Mood:} Monsoon trade, highland fortresses, spirit-ridden passes, oral law.

\textbf{Starting Location:} A rock-chapel at the head of the Silver Pass, where the air is cold and the wind smells of snow. A Ridge-Watcher captain blows a horn of ibex horn: ``The Crown Council has declared a muster. The passes are closing to all foreign traffic. You have until dusk to declare your business – or turn back. The \textit{mizimu} are restless. They taste fear on the wind.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Taharka grows rich on monsoon trade – spices, coffee, and frankincense that travel east to the distant ocean and west to the Gulf. The kingdom has never been conquered; its passes are a graveyard for Ashaani armies and a barrier that Oshiiran grain barges cannot cross. The highlands are a fortress of rock and thin air, where every valley speaks a different dialect and every ridge has a watchtower. On the coast, the monsoon winds bring dhows from Dhahara, Ayokha, and the eastern isles, and the coastal tongue – \textit{Kimvita} – has grown from a dozen voices, the language of the wharves where a merchant can bargain in verse and a captain can curse in seven accents.

The old spirits are not a single kind. \textit{Mizimu} watch from the cairns, demanding tobacco and silence. \textit{Kichocheo} live in baobabs – they throw stones at anyone who counts to nine beneath their branches. \textit{Pepo} ride the monsoon winds; they speak through the mad, and their price is always a white turban left at a crossroads. In the deepest passes, the \textit{popobawa} hunts at the dark of the moon – but only if you whispered its name in doubt the night before. And the passes themselves are alive, breathing with the memory of every army that froze or fell.

\vspace{2mm}
\textbf{Neighbors and Trade Partners:}
\begin{itemize}
  \item \textit{North – Ashaan:} The Veiled Sisters' agents slip through disguised as spice merchants, but the high passes have never fallen to Ashaani armies. The Ridge-Watchers test every stranger with the coffee ceremony – three named cups: \textit{Abol} (welcome), \textit{Tona} (negotiation), \textit{Baraka} (contract) – and now a fourth, the silence cup.
  \item \textit{South – Ngomebe:} The walking cities of the Stone Men are distant neighbors, linked by the Salt Scar and the Aeler engineers who maintain both bearing-foundries and pass-capstans. Ngomebe masons occasionally travel north to study Taharkan rock-chapels.
  \item \textit{East – Dhahara and Ayokha (Monsoon Circuit):} Taharkan coffee and spices ride the monsoon winds east to the Dhaharan coast and the Ayokhan river markets. The \textit{mchango} – a rotating mutual-aid treasury – binds captains across three kingdoms; a Taharkan spice warehouse in Ayokha is considered neutral ground.
  \item \textit{Southeast – The Aelaerem of the Southern Hills:} Pygmy halflings who tend the high passes and trade coffee for stories. Their ninth-cup rite is respected by every Ridge-Watcher.
\end{itemize}

\vspace{2mm}
\textbf{What you notice first:} The smell of roasting coffee and frankincense. The way the air thins as you climb, until each breath is a measured gift. The constant lowing of pack mules on terraced paths.

\textbf{The one rule that kills newcomers:} Never refuse the second cup of coffee. The first (\textit{Abol}) is welcome; the second (\textit{Tona}) is a test of honour; the third (\textit{Baraka}) is a contract. Refuse the second, and you have insulted the house – and the house's ancestors, who listen through the floorboards. And never, ever pour the fourth cup – that is for the \textit{mzimu} of the hearth, and they are always thirsty.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Pass}
Power here is measured in passes, horse lineages, and the favour of the monsoon. The Crown Council rules from the high plateau, but each valley has its own lord, each pass its own toll-keeper. The Ridge-Watchers are the eyes of the kingdom, stationed in rock-chapels that double as signal towers. The horse is not a mount; it is a lineage, a treaty, a weapon. To control the highland breeds is to control the passes – and to control the passes is to remain free. The lowland coast operates under a different law: the \textit{mchango} of the dhows, a rotating mutual treasury where every captain owes a share of every voyage to the community. A captain who cheats the \textit{mchango} finds his ship's name struck from the \textit{laha ya meli} (ship register), and the \textit{mng'aro} of the wharf will point at him every dusk until he pays.

\subsection*{Prominent NPCs of Taharka}
\label{sec:taharka-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Ridge-Watcher Captain Anwar & Commander of the Silver Pass garrison & Keep the passes closed to Ashaani spies & He has not descended below the snow line in twenty years. His shadow is a man who died on the pass. \\
\hline
Merchant-Queen Safiya & Spice House matriarch of the western coast & Control the monsoon trade before the Crown taxes it & She speaks the coastal tongue with an accent that no one can place. Her left eye is a cowrie shell. \\
\hline
Horse Lord Tamrat & Breeder of the royal stables & Sell a foal to the Dhaharan court for a pass-right & He names his horses after dead poets. The horses recite verses in their gait – riddles called \textit{fumbo}. \\
\hline
Sister Hana & Covenant judge of the highland circuit & Resolve disputes between lords without drawing Crown attention & She carries a water-clock that runs on melted snow. It never freezes. \\
\hline
The Spirit-Healer of the Monsoon & Crown advisor, \textit{mganga wa pepo} & Predict the first rains to set the trade season & He has not seen the sun in thirty years. He keeps a \textit{luba} (spirit drum) that plays itself at midnight. \\
\hline
The Pass Spirit (\textit{Kinyamkera}) & A spirit of a soldier who froze on watch & Warn travellers of avalanches and ambushes & It appears as a column of steam. It points with a \textit{kiwiko} (a wooden spoon). \\
\hline
Mama Zawadi & \textit{Buna-tcha} (coffee woman) and spirit medium & Speak for the spirits of the baobab groves & She never walks on bare earth. Her sandals are woven from the roots of a sacred fig. \\
\hline
\end{longtable}

\subsection*{Names of Taharka}
\label{sec:taharka-names}

Inspired by the highland plateau and the coast: melodic, with doubled consonants, honorific prefixes (Mwalimu, Bibi, Bwana), and surnames that denote a clan, a pass, or a horse-line.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Tamrat & Safiya & Berhanu & \textit{the Ridge-Born} \\
\hline
Anwar & Hana & Tesfaye & \textit{the Horse-Tongue} \\
\hline
Dawit & Zawadi & Gebre & \textit{the Monsoon-Watcher} \\
\hline
Kassim & Lij & Mekonnen & \textit{the Ninth Pass} \\
\hline
Solomon & Tsion & Desta & \textit{the Coffee-Blade} \\
\hline
Jabir & Mwanaidi & Asfaw & \textit{the Thin-Air} \\
\hline
Omari & Bahati & Ngoma & \textit{Drum-Voice} \\
\hline
\end{longtable}

\paragraph*{(Farm/Chapel/Pass)} Terraced coffee farm on stone steps; rock-chapel where walls drip condensation; high pass where snow lingers even in summer.

\section*{Spades — Places}
\label{sec:taharka-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Terraced Farm (Coffee)} --- Shrubs planted on stone-step shelves.
  \hfill \textit{Each step is a different family's share – paid in beans, not coin. The \textit{buna-gada} terrace is harvested only once every eight years, at the turn of the \textit{gada} cycle.} `[LABOR]`
\item \textbf{Rock-Chapel (Dripping)} --- A cave-temple where the walls weep condensation.
  \hfill \textit{Drink from the eighth stone, not the ninth. The water tastes of iron and old bargains. A \textit{kichocheo} lives in the drip.} `[RITUAL]`
\item \textbf{High Pass (Summer Snow)} --- A saddle between peaks where the wind never stops.
  \hfill \textit{The snow is blue – the frozen breath of the dead. The \textit{kinyamkera} walks here at dusk, pointing with its spoon.} `[WEATHER]`
\item \textbf{King's Stable} --- A long building of fitted stone, horses in brass halters.
  \hfill \textit{Each stall has a nameplate – a defeated enemy's shield. To bond with a horse, you must answer its \textit{fumbo} (riddle).} `[LEGEND]`
\item \textbf{Spice Warehouse (Coast)} --- A wood-and-coral building that smells of cardamom.
  \hfill \textit{The floor is layered with spices that fell from broken sacks. The ghosts of porters sort them at night – they accept no wages, only the rustle of a \textit{kanga} cloth.} `[TRADE]`
\item \textbf{Monsoon Wharf} --- A stone pier that is underwater half the year.
  \hfill \textit{Each tide mark is a year when a ship never returned. The sailors' widows add new marks with their tears. The \textit{mng'aro} of the wharf appears there on moonless nights.} `[DEATH]`
\item \textbf{Crown Council Hall} --- A roundhouse of stone and horn, built into a hillside.
  \hfill \textit{The hall has nine doors. Each door leads to a different valley. The ninth door is sealed. Behind it is a spider that weaves the future – but it only weaves at the turn of the \textit{gada}.} `[AUTHORITY]`
\item \textbf{Horse Cemetery} --- A field of cairns, each topped with a bronze bell.
  \hfill \textit{The bells ring when a foal is born to the same bloodline. The dead horses speak through the bells – their advice is always a \textit{fumbo}.} `[DEATH]`
\item \textbf{Ridge-Watcher Tower} --- A stone needle on a peak, with a signal fire that never goes out.
  \hfill \textit{The fire burns blue when the passes are closed. It burns green when a foreign army is sighted. It burns red when a \textit{popobawa} walks.} `[WARNING]`
\item[J] \textbf{Coastal Tongue School} --- A thatched hut where children learn \textit{Kimvita}.
  \hfill \textit{The teacher is a one-eyed sailor. He signs his lessons with knots, not words. The knots are a map of the monsoon winds.} `[SOCIAL]`
\item[Q] \textbf{The Spirit-Healer's Observatory} --- A roofless tower with a bronze armillary sphere.
  \hfill \textit{The sphere has nine rings. The ninth ring is broken. The \textit{mganga} refuses to fix it – he says the ninth wind has no name.} `[MAGIC]`
\item[K] \textbf{The Pass of the Ninth Army} --- A defile where an Ashaani host froze to the rock.
  \hfill \textit{On cold nights, you can still see their shapes in the ice. Their lips move – they are still counting. If they reach nine, the \textit{popobawa} answers.} `[DEATH]` `[FEAR]`
\item[A] \textbf{The King's Tomb (Hidden)} --- A cave sealed with a single keystone.
  \hfill \textit{The keystone has a single word carved in \textit{Kimvita}: ``Wait.'' The word changes each \textit{gada} season. Only the Ridge-Watchers know the current one.} `[LEGEND]`
\end{enumerate}

\paragraph*{(Trader/Watcher/Envoy)} Horse trader with a copper bell; Ridge-Watcher captain with a bow over his shoulder; Spice House merchant in robes smelling of cardamom.

\section*{Hearts — People \& Factions}
\label{sec:taharka-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Horse Trader} --- Copper bell on his belt, a ledger of bloodlines in his head.
  \hfill \textit{He names his price in coffee, not coin. He knows every \textit{fumbo} of every horse in the valley.} `[TRADE]`
\item \textbf{Ridge-Watcher Captain} --- Bow of horn and sinew, arrows fletched with eagle feathers.
  \hfill \textit{He can draw a bow at full gallop. He has never missed. He says the mountain guides his hand – but he also leaves a pinch of \textit{tumbaku} at every cairn.} `[COMBAT]`
\item \textbf{Spice House Merchant} --- Robes stained with cinnamon, fingers yellow with turmeric.
  \hfill \textit{She can identify a spice's origin by taste. She has been wrong twice. Both times, the ship sank. She no longer tastes alone – she shares the cup with a \textit{mzimu}.} `[TRADE]`
\item \textbf{Crown Council Envoy} --- Careful, neutral, carrying a staff of office carved from a single piece of ebony.
  \hfill \textit{He speaks in questions. ``Would the honoured traveler perhaps consent to...'' The questions are traps. The answers are leverage. The staff is the \textit{bokku} of the \textit{abba gada}.} `[POLITICS]`
\item \textbf{Coffee Farmer} --- Hands stained black from hulling beans, eyes crinkled from sun.
  \hfill \textit{He offers a cup to every passerby. The cup is clay – from a sacred spring. He never drinks first; that would insult the spirit of the bean.} `[COMMUNITY]`
\item \textbf{Monsoon Captain} --- A sailor with a brass ring in one ear and a scar across his throat.
  \hfill \textit{He speaks the coastal tongue with a Dhaharan accent. His ship is named ``Ninth Wind.'' He has never lost a dhow to a storm – the storms are his cousins, because he pays his \textit{mchango} on time.} `[TRAVEL]`
\item \textbf{Pass Guide} --- A hooded figure with a staff tipped with a horse skull.
  \hfill \textit{She never speaks above a whisper. The wind carries her voice exactly as far as needed. She knows every crevasse by the sound it makes when she spits into it – and which ones are actually the breath of a \textit{kinyamkera}.} `[STEALTH]`
\item \textbf{Sister Hana (Covenant Judge)} --- Grey robe, water-clock, a scale that never balances.
  \hfill \textit{She has ruled on nine border disputes. The losing party always apologizes. The winning party always shares their meal with her – and a cup of \textit{Baraka}.} `[JUSTICE]`
\item \textbf{Horse Lord Tamrat} --- Grey beard braided with horsehair, missing two fingers.
  \hfill \textit{The missing fingers were bitten off by a foal. He kept the foal. It is now his favourite mount. Its name is ``Apology'' – and its \textit{fumbo} is unanswerable.} `[SOCIAL]`
\item[J] \textbf{The Spirit-Healer} --- Blind, pale, wrapped in wool. He smells of wax and ozone.
  \hfill \textit{He predicts the monsoon by the pulse of a \textit{luba} drum. The drum has never been wrong. He feeds it his own dreams.} `[MAGIC]`
\item[Q] \textbf{The Pass Spirit (\textit{Kinyamkera})} --- A column of steam in the cold air, shaped like a soldier.
  \hfill \textit{It points with a wooden spoon. The direction indicates the safest route – but the spoon is cold. If you take the spoon, you can ask one question of the dead. Your hand will freeze to it.} `[DEATH]` `[WARNING]`
\item[K] \textbf{The King (Offstage)} --- Never seen, rarely heard, but his name opens every door.
  \hfill \textit{Rumour says he is a horse. Rumour says he is a \textit{pepo}. Rumour says he is a child of nine who never grew up. The Ridge-Watchers do not correct any rumour.} `[AUTHORITY]`
\item[A] \textbf{The Ninth Heir (Exiled)} --- A prince who lost a succession war, now living as a horse-trader.
  \hfill \textit{He carries a seal made of bone. The bone was his father's thumb. He has not used it. He is waiting for the right moment – or the right \textit{fumbo}.} `[LEGEND]`
\end{enumerate}

\paragraph*{(Landslide/Sickness/Assassination)} A landslide blocks the pass (delay 2 segments); horse sickness spreads — all mounts lose 1 Speed; the king is assassinated — passes close to all foreign traffic.

\section*{Clubs — Complications/Threats}
\label{sec:taharka-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Landslide Block} --- A rockslide covers the pass. Delay 2 segments. `[WEATHER]`
  \hfill \textit{The rocks are warm. The landslide was caused by a signal fire on the peak. Someone wanted the pass closed – and they did not leave an offering at the \textit{kichocheo}'s baobab.}
\item \textbf{Horse Sickness} --- A cough spreads through the stables. All mounts lose 1 Speed. `[DISEASE]`
  \hfill \textit{The sickness came from a Dhaharan ship. The captain is a spy. The horses cough in a rhythm that spells a \textit{fumbo} – a warning about the ninth pass.}
\item \textbf{King Assassinated} --- The Crown Council declares a succession crisis. All passes close. `[POLITICS]`
  \hfill \textit{The assassin was a horse. The horse was poisoned. The poison was in the coffee – the \textit{Baraka} cup. The horse is still alive. It will not speak.}
\item \textbf{Ashaani Raiders (Disguised)} --- A caravan of ``merchants'' is actually a Veiled Sister's scout team. `[SPY]`
  \hfill \textit{They wear Taharkan dress but do not remove their shoes indoors. The innkeeper noticed. She poured them a fourth cup – the silence cup – and a \textit{pepo} now follows them.}
\item \textbf{Coffee Blight} --- A fungus attacks the terraces. Prices triple. Starvation looms. `[DISASTER]`
  \hfill \textit{The blight was spread by a trader from the coast. The trader's ship flew no flag. The fungus writes names on the leaves – the names of those who broke the \textit{gada} oath.}
\item \textbf{Monsoon Delay} --- The winds shift. The first ships are three weeks late. `[WEATHER]`
  \hfill \textit{The delay was predicted by the spirit-healer. He warned the Crown. The Crown did not listen – they refused to renew the \textit{mchango} with the eastern fleets. The spirit-healer has stopped drumming.}
\item \textbf{Spice Warehouse Fire} --- A stack of cinnamon ignites. The whole wharf is at risk. `[DISASTER]`
  \hfill \textit{The fire was set by a rival House. The House has a monopoly on pepper. The smoke smells of cloves – but cloves do not burn. A \textit{mng'aro} is laughing in the smoke.}
\item \textbf{Trade Creole Dispute} --- A contract is voided because of a mistranslated word. `[SOCIAL]` `[DEBT]`
  \hfill \textit{The word means ``shipment'' in \textit{Kimvita} and ``hostage'' in \textit{Kiamu}. The merchant knew. The hostage is the party's ally. The \textit{kinyamkera} points at the merchant's house.}
\item \textbf{Ridge-Watcher Betrayal} --- A captain sells the pass schedule to an Ashaani agent. `[SPY]` `[DEBT]`
  \hfill \textit{The captain's daughter is a spirit-broken slave in Galanina. He had no choice. He carved her name into every stone of the pass – and now her \textit{mzimu} walks there at dusk.}
\item[J] \textbf{The Pass Spirit's Warning} --- The \textit{kinyamkera} points at a traveller. The traveller will die before sunset. `[DEATH]` `[OMEN]`
  \hfill \textit{The traveller is a party member. The spirit is pointing at a shadow, not a person. The shadow is not theirs. It belongs to a \textit{popobawa} that has been following them since they doubted its name.}
\item[Q] \textbf{Crown Council Schism} --- The nine lords cannot agree on a successor. Civil war looms. `[POLITICS]`
  \hfill \textit{The ninth lord is a child. The child speaks only the coastal tongue. No one understands. The child's nanny is the only interpreter – and she is not neutral. She is the \textit{abba gada} of the skipped generation.}
\item[K] \textbf{Horse Line Extinction} --- A plague kills the last mare of the royal bloodline. `[DEATH]` `[LEGEND]`
  \hfill \textit{The mare's foal was stolen by a rival kingdom. The foal is in a Dhaharan stable. The mare's ghost runs the passes every night, whinnying a \textit{fumbo} that no one has answered.}
\item[A] \textbf{Spirit Rising (\textit{Mzimu} Uprising)} --- The old ancestors wake from the cairns. They demand a reckoning. `[MAGIC]` `[FEAR]`
  \hfill \textit{A traveller broke a taboo – they counted to nine near a spirit tree. The \textit{mizimu} want a story that ends with eight. Tell one, or pay in a lock of hair and a cup of \textit{tumbaku}.}
\end{enumerate}

\paragraph*{(Bond/Stone/Pass)} Horse-bond – a foal promised (future String); \textit{Jiwe la Mng'aro} (``the gleam-stone'') – crossing a high pass at half toll; \textit{Kikombe cha Buna} (coffee cup) – trade goods at +1 Position for one scene.

\section*{Diamonds — Rewards/Leverage}
\label{sec:taharka-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Horse-Bond (\textit{Kiapo cha Farasi})} --- A future String. The foal has not yet been born. `[COMMUNITY]`
  \hfill \textit{The bond is a braid of horsehair, braided with a \textit{fumbo}. Wear it, and the mare will remember you. When the foal is born, you will be its first rider – if you answer its riddle.}
\item \textbf{Pass-Stone (\textit{Jiwe la Mng'aro})} --- A carved token. Cross any high pass at half toll. `[TRAVEL]`
  \hfill \textit{The stone is a piece of the \textit{kinyamkera}'s spoon. It is cold to the touch. It warms when a storm is coming – and it burns if you have spoken a doubt about the \textit{popobawa}.}
\item \textbf{Coffee Measure (\textit{Kikombe cha Buna}, Royal Seal)} --- A clay cup. Trade goods at +1 Position for one scene. `[TRADE]`
  \hfill \textit{The cup was used by the king. He drank \textit{Baraka} from it once. The crack is from his tooth. He still has the tooth – it is set in a ring that can open the ninth door.}
\item \textbf{Ridge-Watcher's Horn} --- Blow it to summon a guide from the nearest tower. `[TRAVEL]`
  \hfill \textit{The horn is made from an ibex that the captain killed with a single arrow. The ibex's ghost will lead you – if you give it a story that ends with eight.}
\item \textbf{Spice House Letter of Credit} --- 5 Measures of trade value, redeemable at any coast warehouse. `[TRADE]`
  \hfill \textit{The letter is sealed with cinnamon oil. The seal smells of betrayal. It also smells of profit – and the \textit{mchango} treasury will honour it once before the ink dries.}
\item \textbf{Crown Council Verdict} --- A ruling in your favour on any border dispute. `[JUSTICE]`
  \hfill \textit{The verdict is carved on a slate tile. The tile is also a signal flag. Show it to a Ridge-Watcher, and they will argue for you – but they will also ask for a cup of \textit{Tona}.}
\item \textbf{Monsoon Schedule (Spirit-Healer's Copy)} --- Predicts safe sailing days for the next season. `[WEATHER]`
  \hfill \textit{The schedule is drummed into a clay cylinder. The cylinder must be kept in a jar of water. If the water dries, the predictions invert – and the \textit{luba} will play a warning.}
\item \textbf{Coastal Tongue Phrasebook} --- Gain +1 die to all social rolls on the coast for one journey. `[SOCIAL]`
  \hfill \textit{The phrasebook is a string of knots. Each knot is a curse. The curses are useful. The blessings are not – they are just knots tied by the one-eyed sailor's ghost.}
\item \textbf{Horse Lord's Favour} --- One foal from the royal stables, delivered in a year. `[COMMUNITY]`
  \hfill \textit{The favour is a single white hair from the stallion's tail. The stallion is dead. The hair is still warm – and it answers to its \textit{fumbo} only when the wind is from the north.}
\item[J] \textbf{Pass Spirit's Pointing} --- One safe route through any high pass, guaranteed. `[TRAVEL]`
  \hfill \textit{The \textit{kinyamkera} writes the route in frost on a rock. The frost melts at noon. You must memorize it before then – and you must leave the wooden spoon at the cairn.}
\item[Q] \textbf{The Spirit-Healer's Drum (\textit{Luba})} --- A clay drum that pulses with the monsoon's heartbeat. `[MAGIC]`
  \hfill \textit{Once per campaign, drum the \textit{luba} to summon a \textit{pepo}. The \textit{pepo} will answer one question about the weather – but then it will ask a question of its own. The question is always personal.}
\item[K] \textbf{The Ninth Heir's Seal} --- A bone thumbprint. Commands loyalty from the exiled prince's network. `[AUTHORITY]`
  \hfill \textit{The seal is also a key. It opens a lock in the Crown Council Hall. The lock is the ninth door. The door leads to a room with no floor – only a spider's web that weaves the names of those who have broken the \textit{gada} oath.}
\item[A] \textbf{The King's Blessing} --- A royal decree written in the coastal tongue. One national favour. `[LEGEND]`
  \hfill \textit{The decree is written on a horse's hoof. The hoof was the king's favourite mount. The king still weeps when he sees it – and the mount's \textit{mzimu} will carry you one night, but only if you answer its \textit{fumbo} correctly.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:taharka-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2–5 → 4, 6–10 → 6, J/Q/K → 8, A → 10).
\item Diamonds are codified outcomes (bonds, stones, measures) that change position rather than call for a roll.
\item If any A appears, echo \textbf{monsoon-omens} — horses that speak in \textit{fumbo}, coffee that tastes of snow, a pass that closes behind you, a \textit{luba} drum that plays without a hand, a spider that weaves the future into its web.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:taharka-mechanics}

\begin{itemize}
\item \textbf{Ridge-Watcher Signals:} A network of signal towers on every peak. A fire, a mirror, or a horn can send a message across the entire kingdom in a day. The Ridge-Watchers are neutral in Crown politics; they serve only the passes. A character who spends an action to light a signal fire may reduce the next weather hazard by 1 segment, but the fire also attracts \textit{mizimu} (GM gains 1 SB).
\item \textbf{Mchango of the Dhows (Mutual Treasury):} Along the coast, every captain contributes a fixed share to a rotating treasury. A character who contributes to the \textit{mchango} (by paying a share, repairing a dock, or telling a true story) gains a \emph{Community Favour} token. Spend it to reroll a failed social roll with any coastal merchant or captain. Hoarding wealth without contributing, however, gives the GM 1 SB per session as the community closes ranks – and the \textit{mchango} treasurer may send a \textit{pepo} to remind you.
\item \textbf{Spirit Taboos (\textit{Mwiko}):} The old spirits have specific taboos. Breaking a taboo (counting to nine near a \textit{kichocheo}'s baobab, refusing a second cup of coffee, wearing shoes in a shrine) draws their attention. The GM may then spend 2 SB to have a spirit appear and demand a story that ends with the number eight. A good story pacifies the spirit; a poor one curses the offender with \textit{Mwiko} (\textendash1 die on all rolls until they make an offering of honey and blood).
\item \textbf{Horse Lineage as \textit{Fumbo} (Riddle):} Every highland horse belongs to a lineage that traces back to a legendary stallion – and each lineage has a \textit{fumbo}, a riddle. Knowing a horse's \textit{fumbo} gives +1 die on social rolls with its owner. A character who can recite the \textit{fumbo} of a horse from memory may treat the horse as a temporary ally (Cap 2) for one scene. Forging a lineage (inventing a false \textit{fumbo}) is a capital crime – the Crown Council has executed three forgers in the last decade. Their ghosts guard the pass where they died, and their \textit{fumbo} is unanswerable.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Mzimu wa Kwanza} (The First Spirit) – The Witness. \textit{The ancestor who remembers every oath, every betrayal, every cup of \textit{Tona} refused.}
    \item \textbf{Bwana wa Nyasi} (Lord of the Grass) – The Traveler. \textit{The wind that guides the monsoon dhows and the rider lost in the high passes.}
    \item \textbf{Shetani wa Tisa} (The Spirit of Nine) – The Ninth. \textit{The presence that lives in the missing step, the uncounted wave, the breath after eight.}
    \item \textbf{Bibi wa Kahawa} (Grandmother of Coffee) – Inaea. \textit{The web of trade that binds coast to highlands, merchant to farmer, dhow to mule. Her \textit{fumbo} is the answer to every broken promise.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Monsoon Trade (Dhahara/Ayokha):} Taharka's spice and coffee travel east on the monsoon winds, reaching the Dhaharan coast and the Ayokhan river markets. A delay in Taharka means famine in a dozen ports. The dhows are blessed by Bibi wa Kahawa; a captain who breaks the blessing finds his ship becalmed for nine days – and the \textit{mchango} treasury will not lend him a single sail.
  \item \textbf{The Pass Spirit and the Hollow (Eastern Realms):} The \textit{Kinyamkera} is a spirit of the threshold – a border between living and dead. The Aelaerem say it is a cousin to the Hollow, bound by an oath to the first king. If you see it point at you, you have one day to settle your obligations. The Hollow collects the rest – and the \textit{kinyamkera} will take your spoon.
  \item \textbf{Ashaani Spies and the Veiled Sisters (Ashaan):} Ashaani agents often pose as spice merchants. The Ridge-Watchers have learned to test them by offering coffee – a foreigner who does not know the ritual of three cups (\textit{Abol, Tona, Baraka}) reveals themselves. The Veiled Sisters have begun sending agents who have studied the ritual. The Ridge-Watchers are now testing with a fourth cup – the silence cup. Those who drink it are marked, and a \textit{pepo} follows them.
  \item \textbf{The Ninth Heir and the Kalrantha Amulet (Dhahara):} The exiled prince is rumoured to have bought his horse from a Dhaharan smuggler – a horse that carried a certain green glass amulet. The Crown Council wants to know where the amulet is now. The prince says he sold it for a \textit{jiwe la mng'aro} (pass-stone). The pass-stone is now in the hand of the spirit-healer – and its \textit{fumbo} has not been answered.
\end{itemize}

\subsection*{Taharka Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Pass:} Crumble a crust at the highest point of the pass. Ignore the first \emph{Landslide Block} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Horse's Tail:} Tie a single knot in a horse's tail hair. Cancel \emph{Horse Sickness} or take –1 die on your next riding roll (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Coffee:} Whisper a dead Ridge-Watcher's name into a cup of \textit{Abol}. Gain +1 Effect when negotiating passage this scene, but the ancestors will expect an answer – and the cup will be empty when you look down. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:taharka-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{Ridge-Watchers or Crown Council envoys}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{horse guards or spice warehouse clerks}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{pass guides or rival horse traders}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favourable monsoon window}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Local Ridge-Watcher captain or Spice House factor} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{landslide or fog}.
\item \textbf{Foe Appears:} A \textbf{rival horse lord or Ashaani agent} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Ridge-Watchers, Crown guards, or spice house enforcers} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes – \textbf{king assassinated/monsoon fails/pass closes forever}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{Crown Council, the Spirit-Healer, or the \textit{Kinyamkera}} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically – \textbf{avalanche buries the pass/monsoon surf floods the wharf/coffee terraces catch fire}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice – \textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Taharka (Highlands):} Pass stones glow faintly, horse bells ring without wind, snow falls upward, a \textit{fumbo} whispers in your ear.
\item \textbf{Taharka (Coast):} Monsoon winds carry whispers, spice dust forms faces, the coastal tongue sprouts new words unbidden, a \textit{luba} drum plays a single note.
\item \textbf{Taharka (Spirit Law):} A \textit{kichocheo} laughs from a baobab, a coffee cup cracks, a spider weaves a question into its web, the \textit{popobawa} scratches at your door.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Thin Air}.
\item \textbf{Cut the Pass (Presence + Sway):} On a hit, freeze range for one exchange; if a pass-stone is in your possession, +1 die.
\item \textbf{Weather the Monsoon (Resolve + Sail):} On a hit in \emph{Monsoon Delay}, you pick which ships dock; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Horse, Monsoon \& Spirits)}
\begin{itemize}
\item \textbf{The Horse-Line Thief (Nine Bloodlines)} – \textit{Phases}: Foal Thefts $\rightarrow$ Orchestrated Plague $\rightarrow$ Stable Coup. \textit{Cracks}: recover a stolen foal, expose a false \textit{fumbo}, survive a stampede of ghosts.
\item \textbf{The Monsoon-Breaker (Storm That Stalls)} – \textit{Phases}: Delayed Ships $\rightarrow$ Procession of Empty Wharves $\rightarrow$ Famine Baptism of a Merchant. \textit{Cracks}: relic monsoon drum, survivor testimony, the wind turning against them – and a \textit{pepo} whispering the true schedule.
\item \textbf{The Spirit of the Ninth Pass (\textit{Shetani wa Tisa})} – \textit{Phases}: Miscount Whispers $\rightarrow$ Stone Slides $\rightarrow$ Pass Feast. \textit{Cracks}: recite a counting song that ends at eight, offer a honey-and-blood sacrifice at the \textit{kinyamkera}'s cairn, survive a night on the mountain without fire – and without doubting the \textit{popobawa}'s name.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Taharka, power flows through passes, horse lineages, and the favour of the monsoon – and through the \textit{gada} cycle, the eight-year rotation of political authority. The Crown Council rules by consensus, but each valley lord commands his own Ridge-Watchers. The spice merchants of the coast operate independently, paying tariffs to the Crown but answering to no single lord. The \textit{mchango} of the dhows is a parallel economy of mutual aid – a captain who cheats it finds his ship's name erased from every mooring post, and the \textit{mng'aro} of the wharf will point at him every dusk until he pays.

\textbf{For the GM:}
Power in Taharkan society is measured in altitude, access, and the answering of riddles. Patronage often takes the form of pass-stones, horse-bonds, or coffee cups – each tied to a specific valley, bloodline, or trade route. To emphasize the highland spirit-culture:
\begin{itemize}
\item Tie rewards to physical tokens (carved stone, braided hair, clay cups) that can be challenged, stolen, or voided – and that may carry a \textit{fumbo} of their own.
\item Let rival lords issue conflicting pass-rights, forcing players to choose whose path to follow – and whose \textit{mzimu} to offend.
\item Use the rock-chapels, horse cemeteries, and monsoon wharves as arenas for social contests, where hospitality is law, a cup of coffee is a contract, and a \textit{pepo} may be listening.
\item Introduce spirits as recurring obstacles – they are not enemies but creditors of a different kind, demanding stories rather than silver, and always asking: ``Did you count to eight before you spoke?''
\end{itemize}
In Taharkan society, the mountain remembers every step – and the horse remembers every rider. The spirits remember every miscount.
\end{tcolorbox}

\clearpage
\subsection{Aelaerem of the Southern Hills}
\label{subsec:taharka-aelaerem}

\begin{quote}
  \textit{``They are small, but the mountain does not crush them. They walk where the passes forget their names, and the coffee they trade tastes of mist and memory. Never refuse a cup – the first is welcome, the second is a test, and the ninth cup is for the hollow that lives under the hill. They pour eight and leave the ninth empty. So should you.''}
\end{quote}

\paragraph{Quick Reference}
\textbf{Tagline:} \textit{Small hands, old obligations. The hill remembers every footstep.}

\textbf{Genre / Mood:} Pastoral highland folk, hospitality traps, mist-veiled pygmy clans.

\paragraph{Who They Are.} In the high passes of southern Taharka, where the air thins and the coffee terraces climb into the clouds, live the Aelaerem of the Southern Hills – a pygmy halfling people who were ancient in these mountains before the first Taharkan king raised his banner. They keep to the upper slopes, tending terraced groves of wild coffee, shepherding shaggy mountain goats, and weaving red thread through the antlers of the passes. Outsiders call them \textit{hill-kin} or \textit{cloud-tenders}; they call themselves \textit{the remembered}, because every stone on every cairn has a name, and they are the ones who still speak them.

The Aelaerem have no king, no capital. Their courts are the circle of standing stones at a mountain spring; their law is the counting of cups and the tying of knots. They trade coffee, mountain herbs, and old tales with Taharkan caravans, but they do not trust the lowlands – and the lowlands have learned to respect the ninth cup.

\paragraph{What you notice first:} The smell of roasting coffee and damp moss. The way every threshold of their felt and stone huts has a red thread tied at eye level. The silence when a stranger approaches – and the sudden song when a child decides you are welcome.

\textbf{The one rule that kills newcomers:} Never refuse a second cup of coffee. The first is welcome; the second is a test of honour; the third is a contract. Refuse the second, and you have insulted the house – and the house's ancestors, who listen through the floorboards. And never, ever pour the ninth cup.

\paragraph{The Ninth Cup in the Highlands.} Among the Taharkan Aelaerem, the rite of the ninth cup is subtle: at any feast or negotiation, eight cups are poured. The ninth cup is left empty on the table, facing the door. It is for the Hollow – the presence that walks the high passes when the mist rolls down. If a stranger pours the ninth cup, the Hollow will drink, and the stranger will owe a weight. If the stranger drinks the ninth cup themselves, they have invited the Hollow to sit beside them. The Aelaerem will not intervene. The Hollow is patient.

\paragraph{Coffee \& Knots.} Aelaerem coffee is famous across Taharka, but it is never sold for coin. It is bartered for stories, for promises, for red thread that will be tied around a debtor's wrist. A bag of highland beans comes with a knot – a promise recorded in wool. The knot is cut when the promise is kept. Until then, the debtor can feel it tighten when they sleep.

\section*{Spades — Aelaerem Places (d6)}
\label{subsec:aelaerem-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Cloud Terrace} --- Coffee shrubs planted on stone steps lost in mist.
  \hfill \textit{The mist has a voice. It whispers the names of those who have not yet fulfilled their promises. The eighth step is warm; the ninth is cold.} `[MYSTERY]`
\item \textbf{Red Thread Cairn} --- A pile of stones bound with faded wool; prayers tied to antlers.
  \hfill \textit{Each thread is a promise kept. The cairn grows a handspan each year – but the ninth thread is never tied. It is left loose for the Hollow.} `[RITUAL]`
\item \textbf{Spring Circle} --- A ring of standing stones around a cold mountain well.
  \hfill \textit{The water tastes of iron and old bargains. Drink only from the eighth stone. The ninth stone has no cup.} `[JUSTICE]`
\item \textbf{Felt Hut of the Story-Keeper} --- Low, warm, full of smoke and the smell of roasting beans.
  \hfill \textit{The Story‑Keeper has not left in forty years. The hut has not grown colder. She pours eight cups at dusk, and the ninth she leaves for the \textit{kinyamkera}.} `[HOSPITALITY]`
\item \textbf{Goat Walk} --- A vertiginous path where only the sure‑footed pass.
  \hfill \textit{The goats have names. They will tell you the safe route – if you ask politely, and if you answer their \textit{fumbo}.} `[TRAVEL]`
\item \textbf{The Ninth Cairn} --- A mound of stones with no thread, no offering, no name.
  \hfill \textit{The ninth cairn is for promises that can never be kept. Do not touch it. Do not point. If you must leave something, leave a cup of \textit{Abol} – then walk away without looking back.} `[DEATH]`
\end{enumerate}

\section*{Hearts — Aelaerem People (d6)}
\label{subsec:aelaerem-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Coffee Wife} --- The matriarch of a terrace, keeper of the harvest and the promises.
  \hfill \textit{She ties a knot for every sack. She has never lost a knot. She knows the \textit{fumbo} of every bean.} `[AUTHORITY]`
\item \textbf{Goat-Herd} --- A small figure in a patched cloak, always humming.
  \hfill \textit{His goats answer to the hum. They also answer to the names of dead Aelaerem – and they will not cross a threshold where the ninth cup has been poured.} `[COMMUNITY]`
\item \textbf{Thread-Witch} --- A hedgewise woman who weaves red wool and omens.
  \hfill \textit{She can tie a knot that will hold a ghost – or cut one that should have held. She never cuts the ninth knot; that is for the Hollow to loose.} `[MAGIC]`
\item \textbf{Stone-Speaker} --- An elder who reads the names carved on the old cairns.
  \hfill \textit{He has forgotten his own name. The stones remember it for him. He speaks only in eights – never nine.} `[KNOWLEDGE]`
\item \textbf{Mist-Walker} --- A young Aelaerem who guides caravans through the high passes.
  \hfill \textit{She never uses a compass. She reads the taste of the fog – and she knows which patches of mist are actually the breath of a \textit{kinyamkera}.} `[TRAVEL]`
\item \textbf{The Hollow's Witness} --- An ancient, half‑blind figure who sits at the ninth cairn.
  \hfill \textit{She does not speak. She points. The direction she points is the direction the Hollow will walk tonight. If she points at you, pour a cup of \textit{Baraka} on the ground and do not sleep.} `[DEATH]`
\end{enumerate}

\section*{Clubs — Aelaerem Complications (d6)}
\label{subsec:aelaerem-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Road-Bread Refused} --- A caravan master sneers at the offer of a shared cup.
  \hfill \textit{The pass closes behind him. The mist has teeth – and the \textit{kichocheo} in the nearest baobab is throwing stones.} `[HOSPITALITY]`
\item \textbf{Ninth Cup Poured (By Mistake)} --- A stranger pours the empty cup.
  \hfill \textit{The Hollow drinks. The stranger now owes a weight to a creditor who does not accept silver – only a story that ends with eight.} `[DEBT]`
\item \textbf{Red Thread Cut} --- Someone cuts the thread around a debtor's wrist.
  \hfill \textit{The promise does not vanish. It transfers. The cutter is now the debtor – and the thread will retie itself around the cutter's neck by dawn.} `[JUSTICE]`
\item \textbf{Goat Stampede} --- A flock of shaggy mountain goats barrels through the caravan.
  \hfill \textit{They are not angry. They are fleeing something worse. The mist behind them has a shape – and the shape is counting to eight.} `[FEAR]`
\item \textbf{Coffee Blight} --- A fungus turns the berries black; the terrace sings at night.
  \hfill \textit{The song is a name. The name is the Coffee Wife's. She will not speak of it – but she pours a cup of \textit{Tona} on the ground every night, and the blight retreats an inch.} `[CURSE]`
\item \textbf{Stone-Speaker's Silence} --- The old man stops reading the cairns.
  \hfill \textit{The stones are angry. A promise was broken – a \textit{fumbo} left unanswered. The party is the closest witness. The stones will ask a riddle; if you cannot answer, the ninth cairn will open.} `[JUSTICE]`
\end{enumerate}

\section*{Diamonds — Aelaerem Rewards (d6)}
\label{subsec:aelaerem-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Red Thread Knot} --- A knot tied by the Coffee Wife. Present it to claim a minor promise from any Aelaerem.
  \hfill \textit{The knot tightens when the debtor lies. It loosens when they speak true – but it will never untie completely. That is the interest.} `[SOCIAL]`
\item \textbf{Highland Bean Sack} --- A bag of coffee that never runs dry.
  \hfill \textit{Brew it to gain +1 die on negotiation rolls with any highlander for one scene. The beans smell of mist – and of the \textit{fumbo} of the horse that carried them.} `[TRADE]`
\item \textbf{Moss-Hide Cloak} --- Woven from mountain moss and goat wool.
  \hfill \textit{When standing still in mist, you are invisible to mundane sight. The mist does not betray you – unless you owe a promise. Then it will part, and the \textit{kinyamkera} will point at you.} `[STEALTH]`
\item \textbf{Goat's Tooth Whistle} --- Blow it to call a single sure‑footed goat.
  \hfill \textit{The goat will carry a light load and never slip. It will also judge your character and report to the herd – and the herd's \textit{fumbo} is older than the mountain.} `[TRAVEL]`
\item \textbf{Stone-Speaker's Name} --- A single true name of an ancestor carved on a pebble.
  \hfill \textit{Speak it at a cairn to ask one question of the dead. The answer is always a riddle – and the riddle will cost you a memory if you cannot solve it.} `[KNOWLEDGE]`
\item \textbf{The Hollow's Cup} --- An empty ceramic cup, warm to the touch.
  \hfill \textit{Set it on a threshold, and the Hollow will not cross for one night. After that, the Hollow will remember the favour – and it will ask for a cup of \textit{Abol} at the next new moon.} `[PROTECTION]`
\end{enumerate}

\paragraph{Adventure Seed: The Ninth Cairn's Promise}
A Stone‑Speaker has stopped speaking. The ninth cairn is weeping water that tastes of salt and old blood. The Coffee Wife says a promise was called in – a pledge made by a Taharkan prince to an Aelaerem chieftain, three generations ago. The prince's descendant is in the party's caravan, travelling in disguise. The Hollow walks the passes. The party must either:
\begin{itemize}
\item Find the original promise (hidden in a cairn under a stone that has not been moved in a century – and guarded by a \textit{kichocheo} that demands a story ending with eight).
\item Pay the weight in memory (a cherished recollection of a kindness, given to the Hollow at the ninth cairn – and the Hollow will also take the memory of your face).
\item Or convince the Aelaerem to forgive the promise – which requires a story that moves the stones themselves, and a cup of \textit{Baraka} shared with the \textit{mizimu}.
\end{itemize}
The reward is the Coffee Wife's blessing: a single red thread that can tie any two willing parties into a binding promise, enforceable by the mountain itself – and a \textit{fumbo} that answers itself.

\index{Aelaerem!Taharka highlands}
\index{Coffee ritual!ninth cup}
\index{Aelaerem of the Southern Hills}
\index{Red thread}
\index{Hollow!mountain passes}