\section{Theona --- ``Three Greens, No Ninth''}
\label{chap:theona}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The Green Host rides for those who count to nine. Viterra counts to nine. The Everflame burns the ninth.}

\begin{quote}
  \textbf{Abbess Caelia of Whiterick (Everflame):}  
  ``The Viterran sheriffs count heads for tax. The Moot-keepers count steps for safety. I count souls for salvation. Nine is a sin. Eight is a compromise. Only the Flame is whole.''
\end{quote}

\begin{quote}
  \textbf{Matron of Wells (Theona):}  
  ``The wells here hold more than water---they hold the dreams and the breath of generations. The Viterran sheriffs call it superstition. The Abbess calls it heresy. The Green Host calls it trespass.''
\end{quote}

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The isles of Theona rise green and jagged from the Dolmis, their basalt cliffs wrapped in mists and moorland song. Nominally a duchy sworn to Viterra, in truth the Three Greens heed no crown but their own moots, no law but their taboos. The greatest of these is the \textbf{silence of Nine}: no bell tolls a ninth, no bridge counts that span, no oath dares its syllable. It is said that those who break the ban invite the \textbf{Green Host}---the \textit{Aos S\'{\i}}, the elder folk who were here before the first Viterran longship. 

The Viterran crown claims the isles by right of conquest and papal bull. Its sheriffs impose common law, its bishops raise stone steeples. But the \textbf{Church of the Everflame} watches from Whiterick Abbey. The Abbess schemes to preserve power by pitting Viterra, the native factions, and raiders against each other, seeking to purge the ``pagan'' taboos and impose imperial orthodoxy. To break a taboo is to invite not only the fae but the inquisitor. The sheriff leaves a noose. The Host leaves a hollow in the mist where your name used to be. The Abbess leaves a brand that burns until confession.

\vspace{2mm}
\textbf{What you notice first:} The smell of wet peat and salt. The silence that follows every eighth bell-ring. The hawthorn and blackthorn trees at every crossroads---the \textit{fairy thorns} that no local will touch, and that Viterran surveyors learn to skirt after the first one disappears.

\textbf{The one rule that kills newcomers:} Never pour the ninth cup at a feast. Never speak the ninth name in a bargain. Never step where a step is missing. The Green Host rides for those who complete what should remain incomplete. The Viterran sheriff will hang you for the same offense---but the Host takes your name, not your neck. The Everflame Inquisitor will burn you for the silence itself.
\end{tcolorbox}

\subsection*{The Grey Liturgy: A Syncretic Faith of Theona}
\label{subsec:theona-grey-liturgy}

Deep in the mist-shrouded valleys and along the tide-cut stairs of Theona, a different faith endures---one that does not flee the Ninth but approaches it with ritual, witness, and the patient weight of carried secrets. The \textbf{Grey Liturgy} is not a foreign import. It is the old faith that was nearly erased, pushed to the fens, the barrows, the places where the ninth step is already half-taken. Its practitioners are not priests. They are \textit{keeners} (who sing the dead across), \textit{well-walkers} (who tend the threshold-wells), and \textit{taboo-breakers} (who have taken the ninth step and returned---changed, but alive).

Where Theona's folk practice \textit{avoids} the Ninth, the Grey Liturgy \textit{engages} it---ritually, carefully, with witnesses and salt and the Shepherd's lantern. They are not heretics. They are the ones who remember what the Ninth actually is: not a demon, not a curse, but a door that was left unlocked. The Grey Liturgy carries the key.

\subsubsection*{The Saints of the Threshold (Syncretic Patrons)}

The Grey Liturgy teaches that before the True Light came to \'Eire, there were the \textit{Naomh na Tairs\'{\i}}---the \textbf{Saints of the Threshold}. They were not gods. They were the first to cross from death to life and return. They guide the living, carry the dead, and keep the doors between worlds from swinging shut. The Ninth (the \textit{Caillte}) is not a demon---it is a \textbf{saint who forgot his own name}. He cannot guide because he cannot be called. The Liturgy's highest calling is to remember him.

\begin{longtable}{|p{0.3\textwidth}|p{0.25\textwidth}|p{0.35\textwidth}|}
\hline
\textbf{Patron (Original)} & \textbf{Irish Saint Name} & \textbf{Domain} \\
\hline
The Pale Shepherd & \textit{Naomh Aodh\'an an Tr\'eada\'{\i}} (Saint Aodh\'an the Herdsman) & Passage, death-guiding, the keening \\
\hline
The Confessor Beneath the Bell & \textit{Naomh Lasair\'{\i}ocht na gClog} (Saint Lasair\'{\i}ocht of the Bells) & Burden, sin-eating, confession \\
\hline
The Unbroken Way & \textit{Naomh Ciar\'an na Cothroime} (Saint Ciar\'an of Balance) & Stillness, meditation, threshold-walking \\
\hline
The Witness & \textit{Naomh F\'{\i}onntan an Leabhair} (Saint F\'{\i}onntan of the Book) & Truth, record-keeping, testimony \\
\hline
The Ninth (Lost) & \textit{An Caillte} (The Lost One) & The forgotten threshold, the unnamed door \\
\hline
Ikasha (optional) & \textit{Naomh Sc\'athach na Seift} (Saint Sc\'athach of the Cunning) & Shadow, secrets, the unseen path \\
\hline
The Silent Choir (optional) & \textit{Naomh Maolmhuire na dTost} (Saint Maolmhuire of Silences) & Restraint, protective omission \\
\hline
\end{longtable}

\subsubsection*{The Clergy (Irish Titles)}

\begin{longtable}{|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|p{0.25\textwidth}|}
\hline
\textbf{Role} & \textbf{Irish Title} & \textbf{Meaning} & \textbf{Patron Saint} \\
\hline
Grey Walker & \textit{Tr\'eada\'{\i}} & Herdsman & Naomh Aodh\'an an Tr\'eada\'{\i} \\
\hline
Confessor & \textit{Clogaire} & Bell-Ringer & Naomh Lasair\'{\i}ocht na gClog \\
\hline
Well-Walker & \textit{Cothrom\'oir} & Balancer & Naomh Ciar\'an na Cothroime \\
\hline
Truth-Teller & \textit{Scr\'{\i}obha\'{\i}} & Scribe & Naomh F\'{\i}onntan an Leabhair \\
\hline
Taboo-Breaker & \textit{Caillteoir} & Loser (one who is lost) & An Caillte (no saint) \\
\hline
Shadow-Walker (optional) & \textit{Sc\'ath\'anach} & Shadow-caster & Naomh Sc\'athach na Seift \\
\hline
\end{longtable}

\subsubsection*{Practices of the Grey Liturgy}

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.4\textwidth}|}
\hline
\textbf{Practice} & \textbf{Irish Name} & \textbf{Description} \\
\hline
The Ninth Cup Ritual & \textit{Deoch na gCuig} (Cup of the Ninth) & Poured at the dark moon onto a grave. If the earth drinks, the dead are at peace. If the water pools, the dead are restless. \\
\hline
The Confession of the Dying & \textit{Faoistin an Bh\'ais} (Confession of Death) & The dying speak every secret. The Confessor listens and rings a small iron bell once. The weight transfers. \\
\hline
The Well-Walk & \textit{Si\'uil\'oid an Tobair} (Walk of the Well) & Descent into a sacred well until dawn. If you emerge, you have seen the Ninth and returned. \\
\hline
The Keening & \textit{An Caoineadh} (The Weeping) & A wordless melody that guides the dead. The Grey Walker sings; the dead follow. \\
\hline
The Grey Thread & \textit{Sn\'aithe Liath} (Grey Thread) & Tied around the wrist of the dying; cut when they pass. The thread is burned at the Ninth Cup Ritual. \\
\hline
The Bell-Shrine & \textit{Cloch\'an na gClog} (Stone of the Bells) & A small stone hut where a bell hangs. Confessors ring it when a burden is transferred. \\
\hline
\end{longtable}

\subsubsection*{Saint Aodh\'an the Herdsman (The Pale Shepherd)}

Aodh\'an was a herdsman who lost his flock in a blizzard. He searched for three days. On the third night, he saw a light moving through the snow---a lantern carried by a figure in grey. He followed the light to a cave. Inside, his flock was waiting, warm and unharmed. The figure in grey said: \textit{``You have found them. Now you will guide those who cannot find their own way.''} Aodh\'an became the first \textit{Tr\'eada\'{\i}}. His lantern is the Shepherd's Lamp, carried by Grey Walkers. His feast day is the winter solstice.

\textbf{Symbol:} A lantern with a single grey flame.

\textbf{Invocation:} \textit{``Aodh\'an, who found what was lost, guide this soul to the green hills. The snow is deep. The night is long. But your lamp burns.''}

\subsubsection*{Saint Lasair\'{\i}ocht of the Bells (The Confessor Beneath the Bell)}

Lasair\'{\i}ocht was a nun who tended the dying in a plague village. She heard so many confessions that she began to carry the weight of the dead---their sins, their secrets, their unfinished business. One night, she rang the chapel bell at midnight. The bell cracked. The weight lifted from her shoulders and fell into the earth. A voice from the bell said: \textit{``You have carried enough. Now you will be the one who carries for others.''} Lasair\'{\i}ocht is the patron of confessors. Her bell is the \textit{Clog na Faoistine} (Bell of Confession). Her feast day is the first of November---the day after Samhain, when the dead are closest.

\textbf{Symbol:} A cracked iron bell.

\textbf{Invocation:} \textit{``Lasair\'{\i}ocht, who rang the bell at midnight, take this burden from my shoulders. Let the earth drink what I cannot swallow. Let the bell ring when I am empty.''}

\subsubsection*{Saint Ciar\'an of Balance (The Unbroken Way)}

Ciar\'an was a monk who spent seven years standing in a river, meditating, while the water flowed around him. He never moved. He never fell. When pilgrims asked how he remained still, he said: \textit{``I do not resist the water. I become it. The water flows through me. I am not still---I am the stillness between drops.''} Ciar\'an is the patron of Well-Walkers. His meditation is taught to all who fear death. His feast day is the spring equinox, when day and night are equal.

\textbf{Symbol:} A standing stone with a spiral carved into it.

\textbf{Invocation:} \textit{``Ciar\'an, who stood in the river, steady my breath. The water flows around me, not through me. I am the stone. I am the still point.''}

\subsubsection*{Saint F\'{\i}onntan of the Book (The Witness)}

F\'{\i}onntan was a scribe who copied every manuscript in his monastery. When the Vikings burned the library, he saved nothing---because he had memorised every word. He spent the rest of his life dictating what he had lost. He never forgave himself for not saving the books. But he never lied about what he remembered. F\'{\i}onntan is the patron of Truth-Tellers. His ledger is not a book of judgment---it is a book of \textbf{testimony}. He records what happened, not what should have happened. His feast day is the autumn equinox.

\textbf{Symbol:} An open book with a single grey thread marking a page.

\textbf{Invocation:} \textit{``F\'{\i}onntan, who remembered what was burned, witness this crossing. Write the name in your ledger. Do not judge. Do not forgive. Only record. The truth is enough.''}

\subsubsection*{An Caillte (The Lost One --- The Ninth)}

An Caillte was once a saint like the others. But he forgot his own name. He wandered into the fog and never returned. Now he is the \textit{Dorcha}---the Darkness. He is not evil. He is \textbf{forgotten}. The Grey Liturgy teaches that if someone can remember his true name, he will return and the Ninth threshold will become a door instead of a wound. But no one remembers. And every year, on the night of the dark moon, he knocks. The Grey Liturgy does not open the door. They wait. They listen. They hope.

\textbf{Symbol:} An empty circle. A threshold with no door.

\textbf{Invocation:} \textit{``An Caillte, whose name is lost, we do not call you. We cannot call you. But if you remember yourself, knock twice. We will open the door.''}

\textbf{Note:} The Liturgy does not invoke An Caillte in the same way as the other saints. He is not called upon---he is \textbf{listened for}. Taboo-Breakers are those who have heard him knock and lived. They are the only ones permitted to approach the Ninth threshold.

\subsubsection*{Integration with Theona Generators}

When using the Grey Liturgy in Theona, add the following sprinkles to the existing generators:

\textbf{Spades (Places) --- Add:}
\begin{itemize}
\item \textbf{Cloch\'an na gClog (Bell-Shrine)} --- A small stone hut on a misty headland. A single iron bell hangs inside, cracked and silent. Confessors ring it when a burden is transferred. The Viterran crown declared it a public nuisance. The bell rings anyway. The Abbess tried to melt it down. The bell rang in the forge. The forge went cold.
\item \textbf{Tobar an Chaillte (Well of the Lost One)} --- A sacred well that no one admits exists. The water is black and still. Taboo-Breakers drink from it before their Well-Walk. The Viterran surveyors cannot find it on any map. The Abbess's clerics cannot find it at all. The Lady Beneath drinks from it.
\end{itemize}

\textbf{Hearts (People) --- Add:}
\begin{itemize}
\item \textbf{Clogaire (Confessor)} --- A bell-ringer who carries the weight of the dead. Her hands are scarred from ringing. She never speaks of what she has heard. The Viterran sheriff tried to compel her testimony. She rang her bell. He withdrew the charge. The Abbess tried to buy her silence. She rang her bell. The Abbess's gold turned to ash.
\item \textbf{Tr\'eada\'{\i} (Grey Walker)} --- A guide for the dying. He carries a lantern with a single grey flame. He sings the keening at deathbeds. The Viterran bishop declared his songs ``unlicensed liturgy.'' The songs continue. The Inquisitor declared them heretical. The Inquisitor died. The Tr\'eada\'{\i} sang him across.
\end{itemize}

\textbf{Clubs (Complications) --- Add:}
\begin{itemize}
\item \textbf{The Unsung Bell (1 SB)} --- A confession is not enough. The Confessor cannot ring her bell. The weight must be carried until a witness volunteers.
\item \textbf{The Ninth Knock (2 SB)} --- An Caillte knocks at the threshold. A door that was closed now rattles. The Grey Liturgy waits. The Green Host listens.
\item \textbf{The Burdened Confessor (3 SB)} --- A Confessor collapses under the weight of unspoken sins. She cannot ring. She cannot speak. She can only wait for someone to carry her.
\end{itemize}

\textbf{Diamonds (Rewards) --- Add:}
\begin{itemize}
\item \textbf{Bell-Ringer's Blessing} --- Once, you may ignore a Condition caused by guilt or despair. The Confessor rings her bell for you. The sound is flat, mournful, and final. The Viterran sheriff who received a blessing now sleeps through the night. The Abbess who received one now dreams only of bells.
\item \textbf{Grey Thread} --- A grey thread tied around your wrist. Cut it to transfer a burden to a willing Confessor. The thread cannot be cut by the same hand that tied it. The Viterran governor who cut his own thread now carries his own weight. He is not sleeping. The Abbess who cut a thread for another now sleeps for both.
\end{itemize}

\subsection*{Regional Mood: Three Greens, No Ninth}

Here, peace is a ritual, and every word sharpens the blade that seals it. Theona is claimed by Viterra but ruled by moots, well-keepers, and the old faith. The Viterran sheriffs count heads for the tax-roll; the Brehon judges measure truth by the witness of cattle and the memory of old words. The Viterran bishops raise hymns to the White Flame; the well-keepers leave milk and butter at the thorn crosses. The Everflame Inquisitors count sins in the silence between prayers. Every hedge is a boundary, every bridge a negotiation, every ninth step a door the Green Host might open. And in the mist-shrouded valleys, the Grey Liturgy keeps its candles lit and its bells waiting.

Linn dragon-prows cut through the fog for raiding; Ashaani slavers hunt the coves in stolen coracles; pirates from Zakov and the Vilikari shore trade stolen goods for kelp and salt. Sidhi sailors and \textbf{Aelinnel} factors run cargo up the tide-stairs, and \textbf{Aelaerem} hedge-wives barter red thread for hawthorn passes. To the west, across the narrow sound, the Viterran crown sharpens its taxes and its nooses. Through it all, the Ninth Taboo threads every handshake, every toast, every whispered prayer. To break it is to invite the Green Host---and the Host does not relent. To break Viterran law is to invite the sheriff's noose. To speak the silence is to invite the Everflame's brand. The Theona have learned to break neither---and to bend all three. The Grey Liturgy has learned to listen for the knock.

\noindent\textbf{Starting Location:} A Moot Hill under rain, where vows are carved into standing stone and blackthorn knives gleam beside the ogham pillar. A Bride-Peacemaker draws a blade between two bloodlines. ``The ninth vow was never spoken,'' she says. ``But the Green Host remembers. Bring us the head of the slaver who burned the eastern strand, and the feud will end. Fail, and your name will join the uncounted. Or the sheriff will take you for vagrancy---the crown has declared our moots unlawful assemblies. Or the Abbess will call it heresy. Or, if you listen closely, the Confessor's bell will ring for you.''

\subsection*{Prominent NPCs of Theona}
\label{sec:theona-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Matron Eithne & Keeper of the Sacred Wells & Preserve the old ways against Viterran \& Everflame encroachment & She drinks only from the ninth well---the one no voice names. The crown declared it a nuisance. The Abbess declared it a font of demons. The water still flows, and it tastes of iron and old songs. She has never heard the Confessor's bell. She does not need to. \\
\hline
King Conall & Three-Isles King (elected by the moots) & Walk the narrow causeway between Viterra's law, Everflame's orthodoxy, and the old taboos & He lost his left hand to a Linn raid; the replacement is carved from hawthorn. He signs Viterran writs with the false hand. The Abbess refuses to shake it. The Confessor has offered to carry his lost hand's weight. He has refused. \\
\hline
Br\'{\i}d & Bride-Peacemaker & End feuds before the Green Host rides or the Inquisitors intervene & Her knife has a wooden blade; families cut their palms with wood---blood without iron. She has never lost a client to Viterran justice. The Abbess calls the wood ``unconsecrated.'' She carries a grey thread in her pocket. She does not know why. \\
\hline
Sister Cera & Missionary of the Temple of Light (Viterra) & Convert the islanders to the White Flame through gentle persuasion & She preaches in the old tongue and leaves offerings at hawthorn groves; the locals are not sure whether she is a saint or a heretic. The Viterran bishop calls her a traitor. The Abbess calls her weak. The Confessor calls her patient. \\
\hline
Brother Iarlath & Everflame Inquisitor (Ecktoria) & Purge the `pagan' taboos and impose imperial orthodoxy; serve the Abbess & He carries a censer that never cools, but it smokes green in the fog. He has condemned nine hawthorn sentinels to burning. Eight burned. The ninth vanished overnight. He blames the Host. The Abbess blames him. The Confessor blames the bell. \\
\hline
Abbess Caelia & Abbess of Whiterick (Everflame) & Preserve Everflame power by pitting Viterra, natives, and raiders against each other & She never counts aloud. Her rosary has eight beads. The ninth is a hidden blade. She funds the raiders who harass the Viterran ships. She has heard the Confessor's bell once. She has not slept since. \\
\hline
Sheriff Markham & Viterran crown officer & Stretch the queen's shadow across the Three Greens & He carries a copy of the Statutes of Kilkenny. He has never read them. He uses them to beat witnesses. The Abbess buys his reports before the Governor sees them. He has a grey thread tied around his ankle. He does not remember tying it. \\
\hline
The Lady Beneath & Dream-queen of the hollow hills & Keep the old pacts and claim the ninth name & She has no face; those who see her forget their own names for a year. The Viterran surveyors who saw her now call themselves ``John'' and refuse to speak of it. The Abbess claims she is a demon prince. The Confessor claims she is a bell that has not yet been rung. \\
\hline
Clogaire Malla & Confessor of the Eastern Shore & Carry the weight of the dead who die badly & Her hands are scarred from ringing. She has not rung her bell in seven years. She is waiting. She will not say for what. The Abbess has tried to buy her silence. She rang. The Abbess's gold turned to ash. \\
\hline
Tr\'eada\'{\i} Eamon & Grey Walker of the Fog-Wick Tower & Guide the dying and sing the keening & His lantern has a single grey flame. It has never gone out. He has never lit it. He found it burning. The Viterran bishop declared it a pagan light. He tried to extinguish it. The flame did not notice. The Abbess declared it a demon. The demon did not answer. \\
\hline
\end{longtable}

\subsection*{Names of Theona}
\label{sec:theona-names}

Inspired by the green isles of the western sea: soft consonants, melodic vowels, and surnames that evoke a hill, a well, or a sacred grove.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Conall & Eithne & O'Malloy & \textit{the Ninth} \\
\hline
Cormac & Br\'{\i}d & Mac Gowan & \textit{Well-Sworn} \\
\hline
Iarlath & Cera & \'{O} hAodha & \textit{Hawthorn} \\
\hline
Turlough & R\'{\i}ona & \'{O} Braon\'{a}in & \textit{the Uncounted} \\
\hline
Lorc\'{a}n & M\'{a}ire & \'{O} D\'{a}laigh & \textit{Fog-Born} \\
\hline
Fin\'{a}n & Gobnait & \'{O} Cuilinn & \textit{Ash-Finger} \\
\hline
Eamon & Malla & Mac an Bhaird & \textit{Bell-Ringer} \\
\hline
\end{longtable}

\paragraph*{(Ringfort/Well/Cliff)} Basalt tide-stairs cut into cove; everyone steps past a ``missing'' rung; Saint's Well under hawthorn hung with ribbons.

\section*{Spades --- Places}
\label{sec:theona-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Tide-Stairs} --- Basalt steps into cove; everyone skips the ``missing'' rung.
  \hfill \textit{The missing step is still there---you just can't see it. Step where it should be, or don't step at all. The first Viterran engineer counted nine rungs aloud. He is still counting, somewhere in the fog. The Abbess tried to fill the gap with stone. The stone sank. The Confessor tied a grey thread to the missing step. The thread is still there.}
\item \textbf{Saint's Well} --- Under hawthorn hung with ribbons and secrets; wishes tied in cloth.
  \hfill \textit{The ribbons are knotted in eights. The ninth knot is never tied---only imagined. The crown calls it idolatry. The Abbess calls it witchcraft. The water tastes of iron and old tongues. The Confessors drink from it before they ring.}
\item \textbf{Fog-Wick Tower} --- Burning peat; bell plays eight notes, never nine.
  \hfill \textit{The bell-ringer wears a mask. Underneath, he has no mouth---only a whisper. The sheriff has tried to arrest him three times. Each time, the tower was empty. The Inquisitor tried to silence the bell. The bell rang in his head. The Tr\'eada\'{\i} lights his lantern at the tower. He never speaks to the ringer. They understand each other.}
\item \textbf{Black Bog Causeway} --- White stones like teeth; will-lights test your stride.
  \hfill \textit{The stones were set by a king who drowned. He counts each footstep from below. The Viterran surveyors replaced the stones with wooden posts. The posts burned. The Abbess claims the bog is consecrated ground. It isn't. The Confessor walks the causeway backward. She says the stones are easier to count in reverse.}
\item \textbf{Cliff Ringfort} --- Over kelp-nets; storm-ladders creak like hymns.
  \hfill \textit{The ladders are hauled up at dusk. If you're left below, you sleep with the tide. The crown calls it a ``fortified dwelling without license.'' The fort has no license. It has never needed one. The Everflame wants to build a chapel there. The chapel slides into the sea. The Grey Liturgy built a bell-shrine on the cliff. The bell rings at dusk. The fort still stands.}
\item \textbf{Barrow Field} --- Standing stones; birds hush at noon, the earth remembers.
  \hfill \textit{The stones have ears. Do not speak a secret you cannot bear to lose. The gorse hums with bees that should not be awake. The Viterran antiquary who tried to excavate one now tends the barrow as a shrine. He does not remember why. The Abbess tried to exorcise the field. The exorcism wrote itself on the stones. The Confessor left a bell on the barrow. The bell rang at midnight. The stones listened.}
\item \textbf{Sea-Cave Harp} --- Tide-ropes sing under headland; echoes bargain for safe passage.
  \hfill \textit{The harp has no player. It plays itself when the drowned want to speak. The crown has posted a notice: ``Unauthorized musical performances prohibited.'' The notices wash away. The Abbess says the music is a siren song. She listens anyway. The Tr\'eada\'{\i} sings the keening with the harp. The drowned listen.}
\item \textbf{Uncounted Bridge} --- Nine-arch causeway; locals tally ``eight-and-one.''
  \hfill \textit{The ninth arch is there---but the locals will deny it. Walk it at your own risk. The Viterran cartographer marked it on his map. His map now shows a bridge to nowhere. The Abbess tried to bless the ninth arch. The blessing curdled. The Confessor rings her bell on the ninth arch. The arch remembers.}
\item \textbf{Ogham Grove} --- Carved pillars around an absent ninth post; runes speak in silence.
  \hfill \textit{The missing post was a king's vow carved in heartwood. The runes still curse his line. The Viterran bishop ordered the grove destroyed. The axes broke. The Abbess ordered it burned. The fire turned green. The Taboo-Breakers read the missing post's absence. They say the runes are still there---you just cannot see them.}
\item[J] \textbf{Coracle Harbor} --- Jammed in a notch; upturned boats roof the sheds.
  \hfill \textit{The coracles are made of hide and guilt. Each one has carried a secret. The harbormaster pays no tax. The crown cannot find him. The Abbess cannot buy him. The Confessor has a coracle. She rows it without oars. The tide moves her.}
\item[Q] \textbf{Green Moot Hill} --- Three Greens' court; benches of turf, salt wind judges.
  \hfill \textit{The wind never stops. It carries the votes of the dead. The Viterran sheriff declared the moot an ``illegal assembly.'' The wind knocked his hat off. He has not returned. The Abbess sent a clerk to record the votes. The clerk recorded his own confession. The Tr\'eada\'{\i} sings the keening at the moot's close. The dead vote.}
\item[K] \textbf{High Hall} --- Whale-rib beams; storm banners sleep, kings are chosen.
  \hfill \textit{The kings are chosen by lot---a knotted cord pulled from a pot. The cord has eight knots. The crown's appointed governor was given the ninth knot. The knot came undone. The Abbess offered a golden cord. It strangled the bearer. The Confessor rings her bell when the king is chosen. The king remembers.}
\item[A] \textbf{Lookout Cliff} --- One vantage where all three isles stand in a single gaze.
  \hfill \textit{From here, you can see the Green Host riding on the horizon---if it wants you to. You can also see the Viterran garrison on the far isle. They do not wave. You can see Whiterick Abbey's smoke. It is always green. You can see the Confessor's bell-shrine. The bell is ringing. You cannot hear it.}
\end{enumerate}

\paragraph*{(Moot/Keeper/Court)} Peat-cutter with steady barrow and steadier gossip; well-keeper who ties ribbons and unknots feuds.

\section*{Hearts --- People \& Factions}
\label{sec:theona-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Peat-Cutter} --- Steady barrow, steadier gossip; knows every hidden path.
  \hfill \textit{Her barrow has a false bottom. She hides keepsakes there---and the tax collector when he comes. He is always grateful. She hides Inquisitors too. They are less grateful. She has a grey thread braided into her hair. She will not say who tied it.}
\item \textbf{Well-Keeper} --- Ties ribbons, unknots feuds; wishes and wisdom in equal measure.
  \hfill \textit{She never ties a ninth ribbon. The ninth wish is always her own. She has wished for the sheriff's departure every day for seven years. He is still here. She is patient. She has wished for the Abbess's silence. The silence is growing. She leaves a ribbon at the Confessor's bell-shrine. The bell rings. The wish is heard.}
\item \textbf{Kelp-Netter} --- Reads rip lines like currents; tide as teacher.
  \hfill \textit{His net has a torn patch. He sewed it with thread from a hanged man's shirt. The hanged man was a horse thief. The net has never torn again. The thread glows in the presence of Everflame steel. He found a grey thread in the net last week. He cut it. It grew back.}
\item \textbf{Harp-Satirist} --- A few bars can end a career or a war; music as a weapon.
  \hfill \textit{Her harp is strung with catgut and grudges. Each string is a name. She has a string for the sheriff. She has not played it yet. She is waiting. She has a new string for the Abbess. It hums when the censer smokes. She played a note for the Confessor. The bell answered. She will not play it again.}
\item \textbf{Wick-Warden} --- Tends fog-lamps, counts in eights; guardian of the silent count.
  \hfill \textit{He counts on his fingers---thumb, finger, finger, finger---and stops. The ninth is for the fog. He taught the Viterran lighthouse keeper to count. The keeper lost his job. He taught the Abbess's novice. The novice lost his tongue. He lights a lamp for the Tr\'eada\'{\i}. The Tr\'eada\'{\i} lights his lantern from it. They have never spoken.}
\item \textbf{Taboo-Witness} --- Geas-keeper who knows which words ruin deals.
  \hfill \textit{She carries a bell with no clapper. She rings it with her tongue. The Viterran magistrate demanded she testify. She rang the bell. He withdrew the charge. The Inquisitor demanded she confess. She rang the bell. He confessed for her. She rings her bell when the Confessor rings hers. The sound is the same.}
\item \textbf{Island Abbot} --- Saint's bell, tolerant smile; faith in the old ways.
  \hfill \textit{His bell rings eight times. The ninth ring is the silence that follows. The Viterran bishop sent a letter of censure. The abbot used it to light his fire. The Abbess sent a letter of conversion. The abbot used it to wrap his cheese. He has a grey thread on his bell's clapper. He does not know who tied it. The bell rings truer now.}
\item \textbf{Green Neighbor} --- Polite emissary from under the hill; borrowed boots, strange customs.
  \hfill \textit{He never steps on a crack. He says the earth has teeth. The Viterran tax assessor stepped on a crack. He now has a limp. The Abbess's envoy stepped on a crack. He now preaches a new gospel of dirt. He carries a grey thread in his pocket. He says it is for counting. He will not count to nine.}
\item \textbf{Coracle-Captain} --- Eight-boat flotilla; swift as rumor, sure as the tide.
  \hfill \textit{Her boats have no names. She says naming a vessel invites the Green Host to count it. The Viterran harbormaster tried to register one. He could not find it. It was right in front of him. The Abbess tried to buy one. The gold turned to kelp. She has a grey thread on each prow. The threads are always damp.}
\item[J] \textbf{Bride-Peacemaker} --- Walks between bloodlines with a silver knife; peace in marriage.
  \hfill \textit{Her knife has a wooden blade. The families cut their palms with wood---blood without iron. The Viterran magistrate declared the ritual ``superstitious.'' The knife is now his letter opener. The Abbess declared it ``unconsecrated.'' The blade cuts her robes. She carries a grey thread in her pocket. She ties it around the hands of the wed. The knot never loosens.}
\item[Q] \textbf{Matron of Wells} --- Queen-mother in truth if not in name; wisdom in water.
  \hfill \textit{She drinks only from the ninth well---the one no one admits exists. The Viterran governor demanded to see it. She showed him a dry hole. He believed her. The Abbess demanded to bless it. The water boiled her hand. She has a grey thread tied around her wrist. She tied it herself. She will not say why.}
\item[K] \textbf{Three-Isles King} --- Elected by the moots; whale-bone throne, weathered eyes.
  \hfill \textit{He lost his left hand to a Linn raid. The replacement is carved from hawthorn. He signs Viterran writs with the false hand. The signatures are legally binding. He is not. The Abbess claims he is a puppet. The strings are roots. He has a grey thread tied around his remaining thumb. The thread is warm.}
\item[A] \textbf{Lady Beneath} --- Unseen queen who bargains in dreams; hawthorn's whisper.
  \hfill \textit{She has no face. Those who see her forget their own names for a year. The Viterran surveyor who saw her now calls himself ``John.'' He does not remember his wife's face. The Abbess claims she is a demon. The demons agree. She has a grey thread wrapped around her missing face. She will not remove it. She is waiting.}
\end{enumerate}

\paragraph*{(Fog/Feud/Taboo)} Ground-mist erases tracks; horns travel, edges do not; bog-lights lure teamster off the causeway.

\section*{Clubs --- Complications/Threats}
\label{sec:theona-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Ground-Mist} --- Erases tracks; horns travel, edges do not; navigation by sound alone.
  \hfill \textit{The mist has a voice. It whispers your name---but only once. The Viterran patrol that entered the mist returned speaking only in questions. The Everflame patrol returned speaking in confessions. The Confessor walks into the mist willingly. She rings her bell. The mist answers. She does not tell anyone what it says.}
\item \textbf{Bog-Lights} --- Lure teamster off the causeway; recovery costs time and pride.
  \hfill \textit{The lights are the souls of those who broke their geas. Follow them, and you'll never find your way back. The Viterran tax collector followed them. He now lives in the bog. He has no use for silver. The Inquisitor followed them. He found a shrine. He burned it. It grew back. The Tr\'eada\'{\i} follows them. He sings the keening. The lights listen. They lead him to the dead.}
\item \textbf{Spoken Geas} --- No ninth word in parley---or the oath curdles like milk.
  \hfill \textit{The word you cannot speak sits on your tongue like a stone. The Viterran merchant who said ``nine'' in a pact now sells only in eights. His customers have not noticed. The Abbess's priest said ``nine'' in a prayer. The prayer answered itself. The Confessor said ``nine'' once. She rang her bell. The word is gone. She cannot remember it.}
\item \textbf{Salt-Rot} --- Smokehouse fails; fish spoil, tempers rise like the tide.
  \hfill \textit{The salt was cut with grave dust. Someone wanted the winter to be hard. The Viterran garrison's supplies rotted. They bought from the locals. The price was a promise. The Everflame pantry rotted. They blamed the Host. The Host blamed the Abbess. The Confessor salted her own fish with ash from her bell-shrine. The fish did not rot. The shrine is warm.}
\item \textbf{Procession Clash} --- Saint-day vs.\ hawthorn taboo; shutters slam, peace fractures.
  \hfill \textit{The saint's banner and the green branch cannot touch. When they do, blood follows. The Viterran bishop ordered the hawthorn removed. The tree grew back overnight. The Abbess ordered the saint removed. The saint stayed. The Abbess left. The Confessor rang her bell between them. The banner and the branch parted. The crowd parted. She walked through. No one followed.}
\item \textbf{Wave-Count} --- Eighth swell, Nameless follows; boats tremble, sailors pray.
  \hfill \textit{The ninth swell never comes. But something rides the eighth. The Viterran customs cutter that chased a smuggling boat did not see the eighth swell. It is at the bottom. The Everflame ship saw it. It turned back. The Tr\'eada\'{\i} sails into the eighth swell. He sings the keening. The Nameless listens. It turns away.}
\item \textbf{Ringfort Wakes} --- A bone-judge is ``consulted''; now you're named in the old law.
  \hfill \textit{The bone-judge's skull is filled with peat smoke. It speaks in riddles---and judgments. The Viterran magistrate who ignored a bone-judge's ruling now has no memory of the case. The Inquisitor who ignored it now has no memory of his faith. The Confessor asked the bone-judge a question. The skull was silent. She rang her bell. The skull spoke. She does not remember the answer.}
\item \textbf{Ninth Law} --- Something vital cannot be counted, measured, or named.
  \hfill \textit{You try to count your crew. The number keeps changing. The Viterran census taker who tried to count the islanders now lives in a cave. He counts seagulls. The Abbess's scribe tried to count the sins. The ink ran out. The Confessor counts her bell-rings. She stops at eight. The ninth ring is silence. The silence counts.}
\item \textbf{Net-Surge} --- Harbor stairs; kelp wraps ankles and alibis like truth.
  \hfill \textit{The kelp moves against the current. Someone underneath is pulling. The Viterran harbormaster who tried to dredge the kelp is now missing. The kelp is thicker. The Everflame divers found a bell. It rang underwater. The Confessor dives for the bell. She rings it. The kelp loosens. The drowned listen.}
\item[J] \textbf{Bride-Theft} --- At Moot---peace flips to feud; you're caught mid-ritual.
  \hfill \textit{The stolen bride's hand is cold. She has been dead for a year. The Viterran sheriff who tried to stop the theft now wears a wedding ring. He does not remember the wedding. The Abbess tried to sanctify the marriage. The ring turned to wood. The Confessor rings her bell at the wedding. The dead bride hears. She does not answer.}
\item[Q] \textbf{Mainland Return} --- Exile with papers; elders arrested ``for progress.''
  \hfill \textit{The papers are stamped with a Viterran seal. The seal is still warm---and wrong. The elder who was arrested now works in the governor's garden. He does not speak. The plants are very healthy. The Abbess offered him sanctuary. He is now her gardener. The plants are very healthy. The Confessor offered him a grey thread. He tied it around his wrist. He sleeps now. The plants are less healthy.}
\item[K] \textbf{Green Host Rides} --- Hooves like drums, cattle scatter, sentries sleep.
  \hfill \textit{The riders have no faces---only masks of woven hawthorn. The Viterran garrison that saw the Host now stands watch at noon. They will not explain why. The Everflame garrison that saw the Host now prays at noon. They pray to the Host. The Tr\'eada\'{\i} watches the Host ride. He lights his lantern. They do not look at him. They are not interested in the living.}
\item[A] \textbf{Great Fog} --- Routes loop, bells lie, every promise takes the long way home.
  \hfill \textit{The fog is old. It remembers vows spoken to the dead and the silence that answered. The Viterran governor who swore to ``pacify'' the isles is still lost in it. His dispatches arrive late. They make no sense. The Abbess's dispatches arrive early. They make too much sense. The Confessor walks into the Great Fog. She rings her bell. The fog parts. She walks back. She does not tell anyone where she went.}
\end{enumerate}

\paragraph*{(Blessing/Token/Right)} Well-blessing for a named ford crossing; moot token with eight-knotted cord skips local queues.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:theona-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Well-Blessing} --- Named ford crosses you safely, once; water remembers your need.
  \hfill \textit{The blessing is a ribbon. Tie it to your wrist. Do not tie a ninth knot. The Viterran sheriff's horse refused to cross the blessed ford. He had to swim. The Abbess's horse crossed. It drowned. The Confessor tied a grey thread to the ribbon. The thread remembers. The ford remembers.}
\item \textbf{Moot Token} --- Eight-knotted cord; skip any local queue or judgment.
  \hfill \textit{The cord is braided from horsehair and silence. Cut a knot to speak. The Viterran magistrate who tried to arrest a token-holder found his courtroom empty. The case was dismissed. The Inquisitor who tried to arrest one found his charges dismissed. By him. The Confessor's token has nine knots. She has not cut any of them. She is waiting.}
\item \textbf{Harbor-Green} --- Painted mark exempts your boat from the next levy; tide as ally.
  \hfill \textit{The mark is a circle with an eighth cut. The missing slice is for the Host. The Viterran customs officer who tried to board a marked boat now gathers shells on the beach. The Everflame officer who tried to board it now preaches to the shells. The Confessor's mark has a grey thread painted through the cut. The tide rises when she rings.}
\item \textbf{Hawthorn Pass} --- Immunity from a single local taboo (for one scene).
  \hfill \textit{The pass is a hawthorn twig. Carry it in your mouth. Do not chew. The Viterran inquisitor who tried to burn a pass-bearer found his pyre would not light. The Abbess tried to bless the twig. It prickled her. The Confessor tied a grey thread to the twig. The thread is wet. The twig is growing.}
\item \textbf{Bell-Right} --- Lawful ring at a fog-wick summons the watchers; law travels on sound.
  \hfill \textit{The bell has no clapper. You must ring it with your own breath. The Viterran constable who heard the bell now refuses to speak of it. His ears are scarred. The Abbess's novice who heard it now speaks only in riddles. The Confessor rings this bell with her breath. The bell rings with her voice. The watchers come. They do not speak.}
\item \textbf{Coracle Share} --- Crew and oars pledged for one crossing; tide as a steed.
  \hfill \textit{The share is a hazel wand. Break it in half to claim your passage. The Viterran governor who tried to seize a coracle share found his hands would not close. The Abbess tried to buy the share. The wand turned to ash. The Confessor's wand has a grey thread tied around it. The thread is damp. The tide knows her name.}
\item \textbf{Ogham Ruling} --- Monk's reading settles a dispute in your favor; runes as law.
  \hfill \textit{The ruling is carved on a slate. The slate is damp. The runes may change by morning. The Viterran judge who ruled against an Ogham ruling now dreams in runes. He cannot read them. The Abbess's judge ruled against it. The slate cracked his gavel. The Confessor's slate has a ninth rune. It appears at dawn. It disappears by noon. No one has read it.}
\item \textbf{Unnumbered Right} --- Ignore the Ninth prohibition for one action (binds you to an answer later).
  \hfill \textit{The right is a silence. Use it, and the Green Host will ask for its name. The Viterran spy who used it now has no shadow. The Host has his name. The Abbess's spy who used it now has no reflection. The Flame has his soul. The Confessor used it once. She rang her bell. The Host did not ask. The bell answered for her.}
\item \textbf{Salt-Cure License} --- Process scarce fish despite ration edicts; preservation as power.
  \hfill \textit{The license is signed with a thumbprint. The thumb is not yours. The Viterran ration officer who refused a licensed boat now has no appetite. He eats only ash. The Abbess's officer who refused it now eats only salt. He is very thirsty. The Confessor's license is signed with a grey thread. The thread is damp. The fish are preserved. The license is warm.}
\item[J] \textbf{Bride-Charter} --- Two feuding houses sheath knives for a season; peace in marriage.
  \hfill \textit{The charter is written on a wedding veil. The bloodstains are from the first cut of peace. The Viterran crown's peace commissioners have tried to copy the charter. Their copies burn. The Abbess tried to consecrate it. The veil caught fire. The Confessor tied a grey thread to the veil. The veil did not burn. The peace held.}
\item[Q] \textbf{Private Moot} --- With Matron and King; whispers become policy.
  \hfill \textit{The moot is held in a fog hollow. What is spoken there is never written. The Viterran governor's spy who tried to listen now speaks only in fog. He is very popular with the locals. The Abbess's spy who tried to listen now speaks only in confessions. He is very popular with the Inquisitors. The Confessor rings her bell at the moot's close. The fog listens. It does not repeat.}
\item[K] \textbf{Whale-Escort} --- Eight-oar guards across the Dolmis reach; strength in numbers.
  \hfill \textit{The crew carries your name on a scrap of vellum. If the paper gets wet, you are alone. The Viterran warship that tried to intercept a whale-escorted vessel now has no name. The Everflame ship that tried to intercept it now has no crew. The Confessor's escort has a grey thread instead of a name. The thread is wet. The crew still knows her.}
\item[A] \textbf{Green Favor} --- One uncanny exception to custom, spoken beneath hawthorn.
  \hfill \textit{The favor is a dream. When you wake, the hawthorn will remember---even if you don't. The Viterran governor who asked for a Green Favor now believes he is a king. The actual king is amused. The Abbess who asked for a Green Favor now believes she is a saint. The Host is amused. The Confessor asked for a Green Favor. She dreamed of a grey thread. She woke with it tied around her wrist. She has not untied it.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:theona-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (blessings/tokens/rights) that change position rather than call for a roll.
\item The ``No Ninth'' custom: whenever a 9 appears in your seed, add an omission to the scene (a missing step, an unsaid name, an unseen guest). If the 9 is a Diamond, you may break the taboo once---but someone will come to collect later.
\item If any A appears, echo \textbf{green and count} motifs---hawthorn that whispers, bells that miscount, promises that take the long way home. If the A is a Heart, a Confessor may appear. If the A is a Spade, a bell-shrine. If the A is a Club, the Green Host rides. If the A is a Diamond, the Grey Liturgy offers a thread.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:theona-mechanics}

\begin{itemize}
\item \textbf{The Silent Count:} Players and GMs are encouraged to omit a detail (a name, a step, a tally) when the Ninth appears, echoing Theona's taboo. This is how the islanders evade Viterran law: they simply do not say what the law requires. The law cannot count what is not spoken. \textbf{Mechanical Effect:} If a player speaks the ninth name/step/item in a scene, tick the \textbf{Green Host Rising} timer by 1. If they omit it successfully, gain 1 Boon. If they tie a grey thread instead of speaking, gain 2 Boons and do not tick the timer.
\item \textbf{Green Host's Omen:} Any dream sequence or omen may be framed as the ride of the Green Host---hoofbeats across the sky, wind through the moors, sea spray tasting of salt and blood. The Viterran garrison has learned to ignore the omens. The garrison is understrength. The Everflame Inquisitors document them as evidence of demonic activity. The Grey Liturgy interprets them as messages from the dead.
\item \textbf{Bride-Peace Feasts:} Weddings, truces, and funerals blur; sharing a feast can end a feud for a season---or doom a guest if the ninth cup is poured. The Viterran crown has outlawed these feasts. The feasts continue. The crown's officers attend. They never pour the ninth cup. The Abbess sends clerics to bless the cups. The cups break. The Confessor pours the ninth cup into a cracked bell. The bell rings. The feast is blessed.
\item \textbf{The Grey Thread:} A character may tie a grey thread around their wrist as a symbol of a burden carried or a promise to a Confessor. The thread cannot be cut by the same hand that tied it. Once per session, the wearer may transfer a Condition caused by guilt, fear, or despair to a willing Confessor NPC, clearing the Condition. The Confessor gains 1 Fatigue. The thread must be burned at a bell-shrine within a week, or the Condition returns.
\item \textbf{The Viterran Plantation Timer [8]:} Tracks the Viterran crown's efforts to colonize the isles. Advances when the party carries the king's salt (guides a patrol, enforces a writ, speaks against a moot) or when a Viterran sheriff successfully imposes common law. Retreats when the party supports a moot, hides a fugitive, or exposes a crown official's corruption. When full, a new plantation is declared: a Viterran town is built, lands are confiscated, and a new sheriff is appointed. The Green Host's shadow grows. The Confessor's bell rings. The dead watch.
\item \textbf{The Everflame Orthodoxy Timer [6]:} Tracks the Abbess of Whiterick's influence. Advances when the party aids the Inquisition, purges ``heresy,'' or accepts Everflame gold. Retreats when the party exposes Everflame manipulation, protects native rites, or turns Everflame allies against the Abbess. When full, the Inquisition establishes a permanent tribunal; native rites become illegal under penalty of burning. The Grey Liturgy goes underground. The bells fall silent.
\item \textbf{Brehon Law vs.\ Common Law:} In any legal dispute, the party may choose to invoke Brehon law (custom, precedent, witness) or Viterran common law (statute, seal, written evidence). Brehon law grants +1 die if the party can produce a respected elder or a witnessed oral history; Viterran law grants +1 die if the party has a writ or a seal. Choosing one over the other shifts the Viterran Plantation timer by 1 segment toward or away from the crown. Choosing to ignore both (relying on Taboo) shifts the Green Host timer. Choosing to invoke the Grey Liturgy (a grey thread and a witnessed silence) shifts none---but the Confessor will remember.
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{Saint Br\'{\i}d of the Unquenched Well (Syncretic):} A local saint venerated by both the Temple of Light and the old well-keepers. Her shrine is a well hung with red ribbons; pilgrims who tie an eighth knot and leave a silver coin may ask one question of the Green Host---but the answer comes in a dream, and the dream always takes a memory. The Viterran bishop has declared her a heretic three times. Each time, the declaration was found burned. The well still flows. The Abbess claims the well is a portal. She is right. The Confessor claims the well is a bell. She rings it. The water answers.
  \item \textbf{Saint Colm\'{a}n of the Eighth Bell (Oath of Mercy and Grace):} A missionary who tried to convert the isles by ringing a ninth bell. The Green Host took his tongue, and he spent the rest of his life speaking in eights. His hospice still stands near the Fog-Wick Tower; it heals the fog-sick, but every patient leaves unable to count past eight. The Viterran crown has tried to close the hospice. The hospice's patients now include the governor's son. The Abbess has tried to recruit him. He speaks only in eights. The Confessor rings a bell for him at dawn. He counts to eight. He stops. He is healing.
  \item \textbf{The Dullahan (The Headless Horseman):} A fae omen that rides when a Viterran sheriff is about to die. It carries its own head under its arm, and it calls the sheriff's name. Those who hear it are dead before dawn. The current sheriff has heard it three times. He is still alive. He is very afraid. The Abbess has heard it once. She laughed. The head did not laugh. The Confessor heard it once. She rang her bell. The head turned. It did not speak.
  \item \textbf{The Banshee (Bean Sidhe):} A spirit who keens when a pure-blooded Theona chieftain dies. The Viterran crown has tried to suppress her songs as ``noise pollution.'' The keening continues. The chieftains die. The crown's record-keepers have learned to stop at eight. The Abbess's record-keepers stop at seven. They fear the eighth. The Tr\'eada\'{\i} sings the keening with her. The chieftain dies well. The banshee is grateful.
  \item \textbf{The Morrigan (The Phantom Queen):} A war goddess who appears as a crow before battle. She does not choose sides. She chooses the dead. The Viterran soldiers who see a crow before a skirmish often desert. The local fighters see a crow and sharpen their blades. The Morrigan is amused. The Abbess's soldiers see a crow and pray. The crow eats their prayers. The Confessor sees a crow and rings her bell. The crow listens. It does not choose. It waits.
  \item \textbf{The Ninth:} In any Theona scene, the ninth person to speak in a room draws the Green Host's attention. The wise hold their tongues until someone else breaks the count. Those who speak ninth are said to see a pale rider on the horizon the next night---and the rider is always one name closer. The Viterran governor who spoke ninth now has a portrait of himself that ages instead of him. He has not looked at it in a year. The Abbess who spoke ninth now has a portrait that does not age. She looks at it every day. The Confessor who spoke ninth rang her bell. The rider passed. She does not have a portrait.
\end{itemize}

\subsection*{Theona Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Bog:} Toss a crust into the mist before crossing a causeway. Ignore the first \emph{Ground-Mist} penalty.
  \item \textbf{Knot the Eighth:} Tie a single knot in a ribbon before a moot. Cancel \emph{Spoken Geas} or take --1 die on your next oath (your choice).
  \item \textbf{Name to the Well:} Whisper a dead enemy's name into Saint's Well. Gain +1 Effect against his kin for the scene, but the Matron will remember your name. The Abbess will also remember.
  \item \textbf{Thread to the Confessor:} Tie a grey thread around your wrist before a dangerous journey. Once, when you would take a Condition from fear or despair, you may transfer it to a Confessor. The Confessor will carry it. The thread will turn black. Burn it at a bell-shrine.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In Theona, power flows through ancient custom, sacred spaces, and the careful observance of taboo. The Three Greens maintain their independence through a web of moots, well-keepers, and geas-keepers who ensure the old laws are honored. The King and Matron rule by consensus and tradition, while the unseen Lady Beneath influences events through dreams and whispers. Against them stands the Viterran crown---a power of sheriffs, bishops, and undertakers who seek to replace Brehon law with common law, hawthorn with church stone, and the Green Host with the White Flame. And watching from Whiterick Abbey, the Everflame Abbess schemes to preserve power by pitting Viterra, the native factions, and raiders against each other.

In the mist, a quieter power waits. The Grey Liturgy does not scheme. It rings. It carries. It listens. Its Confessors hold the weight of the dead. Its Grey Walkers guide the dying across. Its Well-Walkers descend into the Ninth and return. It offers no gold, no land, no writs. It offers only a grey thread and a cracked bell. For those who have lost too much, that is enough.

\textbf{For the GM:}  
Patronage in Theona revolves around blessings, exemptions, and the delicate balance of keeping or breaking taboos. Rewards often take the form of tokens, rights, or divine favor that can be woven into greater standing. To emphasize the colonial conflict:
\begin{itemize}
\item Tie rewards to visible symbols (ribbons, tokens, blessings) that can be challenged, stolen, or voided by crown officials.
\item Let the Viterran crown, the Green Host, and the Everflame Abbess grant conflicting protections---players must choose whose favor matters more, and whose curse is worse.
\item Use the moots, wells, and causeways as arenas for social contests, where knowledge of Brehon law (and its deliberate omissions) determines success against Viterran common law and Everflame orthodoxy.
\item The Viterran Plantation Timer and Everflame Orthodoxy Timer are visible campaign fronts. Players can accelerate or delay them, but once they fill, the isles change forever---new towns, new sheriffs, new taboos, new inquisitors.
\item For parties drawn to the Grey Liturgy, add a third timer: \textbf{The Unringing Bell [8]}. Advances when the party breaks a taboo without offering a grey thread, ignores a deathbed confession, or fails to carry a burden. Retreats when the party ties a grey thread, rings a bell-shrine, or helps a Confessor. When full, the Grey Liturgy withdraws its protection. The Confessors will not ring for you. The dead will not listen.
\end{itemize}
In Theona, your word is your weave, and your weave determines whether you walk the green path or invite the Host---or the sheriff's noose---or the Abbess's brand---or the Confessor's silence.
\end{tcolorbox}

% Index entries
\index{Theona|see{Three Greens, No Ninth}}
\index{Theme generator}
\index{No Ninth generator}
\index{Places!Theona}
\index{People!Theona}
\index{Complications!Theona}
\index{Rewards!Theona}
\index{Matron of Wells}
\index{Three-Isles King}
\index{Lady Beneath}
\index{Saint's Well}
\index{Green Moot Hill}
\index{Black Bog Causeway}
\index{Sea-Cave Harp}
\index{Ogham Grove}
\index{Bride-Peacemaker}
\index{Taboo-Witness}
\index{Green Host}
\index{Hawthorn taboo}
\index{Uncounted Bridge}
\index{Fog-Wick Tower}
\index{Coracle harbor}
\index{Linn raiders}
\index{Ashaani slavers}
\index{Sidhi sailors}
\index{Vilikari traders}
\index{Aelinnel factors}
\index{Aelaerem hedge-wives}
\index{Saint Br\'{\i}d}
\index{Saint Colm\'{a}n}
\index{Covenant of the Counted}
\index{Order of the Unnumbered Cup}
\index{Dullahan}
\index{Banshee}
\index{Morrigan}
\index{Viterran Plantation Timer}
\index{Everflame Orthodoxy Timer}
\index{Abbess of Whiterick}
\index{Brehon law}
\index{Viterran crown}
\index{Church of the Everflame}
\index{Grey Liturgy}
\index{Naomh Aodh\'an an Tr\'eada\'{\i}}
\index{Naomh Lasair\'{\i}ocht na gClog}
\index{Naomh Ciar\'an na Cothroime}
\index{Naomh F\'{\i}onntan an Leabhair}
\index{An Caillte}
\index{Confessor}
\index{Grey Walker}
\index{Well-Walker}
\index{Truth-Teller}
\index{Taboo-Breaker}
\index{The Unringing Bell}

\subsection*{Thematic SB Spend Table}
\label{sec:theona-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{moot-keepers, well-guardians, Viterran sheriffs, or Everflame clerics}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{coracle crews, ringfort sentries, crown patrols, or Inquisitor spies}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{trackers, green neighbors, crown informers, or Host riders}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable tide}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\item \textbf{Unsung Confession (Grey Liturgy):} A Confessor cannot ring her bell. The weight must be carried until a witness volunteers.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Wick-warden, moot official, Viterran bailiff, or Everflame novice} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{mist shift, taboo breach, a sheriff's roadblock, or an Inquisitor's ward}.
\item \textbf{Foe Appears:} A \textbf{feuding house member, bone-judge, crown soldier, or Everflame penitent} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\item \textbf{The Ninth Knock:} An Caillte knocks at a threshold. A door that was closed now rattles. The Grey Liturgy waits. The Green Host listens.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{coracle crews, green neighbors, moot enforcers, crown militia, or Everflame zealots} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes---\textbf{taboo broken/green host rides/feud declared/crown warrant issued/Inquisition summons}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\item \textbf{The Burdened Confessor:} A Confessor collapses under the weight of unspoken sins. She cannot ring. She cannot speak. She can only wait for someone to carry her.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{King, Matron, Lady Beneath, Viterran Governor, or Abbess of Whiterick} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically---\textbf{fog rises/bells ring/causeway shifts/crown forces arrive/Abbess's censer smokes}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice---\textbf{omen appears/blessing lost/curse manifests/Host rides}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\item \textbf{The Unringing Bell:} The Grey Liturgy withdraws its protection. Confessors will not ring. The dead will not listen. The party is alone.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Theona (Silent Count):} Steps disappear mid-climb, names vanish from lips, the ninth item in a list goes unmentioned.
\item \textbf{Theona (Green Host):} Hoofbeats echo without a horse, mist takes familiar shapes, dreams predict the next action.
\item \textbf{Theona (Viterran Law):} A writ is declared void, a sheriff's seal cracks, a witness remembers the ninth name.
\item \textbf{Theona (Everflame Orthodoxy):} A prayer is answered wrongly, a brand burns without fire, a confession is heard by the wrong ears.
\item \textbf{Theona (Grey Liturgy):} A bell rings without a ringer, a grey thread tightens on a wrist, a Confessor's shadow stretches toward the dead.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth):} On a hit, shift one band. On a strong hit, also impose \emph{Fog Bank}.
\item \textbf{Cut the Writ (Presence + Sway):} On a hit, freeze range for one exchange; if a moot token is in your possession, +1 die.
\item \textbf{Weather the Host (Resolve + Spirit):} On a hit in \emph{Green Host Rides}, you pick who is named; on a miss, the GM does.
\item \textbf{Defy the Abbess (Spirit + Deception):} On a hit, hide a taboo act from the Inquisition; on a miss, gain \emph{Marked for Purity}.
\item \textbf{Ring the Bell (Spirit + Lore):} On a hit in \emph{Grey Liturgy}, transfer a Condition to a willing Confessor; on a miss, the Confessor cannot ring---the weight doubles.
\end{itemize}

\subsection*{Boss Seeds (three-phase foes)}
\begin{itemize}
\item \textbf{The Unsung Bell (forbidden note given will)} --- \emph{Phases}: Hum in Mist $\rightarrow$ Echo Under Well-stone $\rightarrow$ Command of Fog. \emph{Cracks}: bind its clapper with eight ribbons; let a rival speak the ninth and pay; drown the echo at the Sea-Cave Harp. \textit{Grey Liturgy variant}: let a Confessor ring it once. The bell will shatter or sing. No one knows which.
\item \textbf{Bride-Thief of the Green Host (dream raider)} --- \emph{Phases}: Invitation in Sleep $\rightarrow$ Procession Across Moor $\rightarrow$ Feast Without Ninth Cup. \emph{Cracks}: cut the dream-veil with silver; sing satire at the Green Moot Hill; trade a Bride-Charter. \textit{Grey Liturgy variant}: offer a grey thread instead of a bride. The Host may accept. The thread will be tied around the King in Green's wrist. He will not cut it.
\item \textbf{The Sheriff's Black Assize (Viterran enforcer)} --- \emph{Phases}: Plantation Decree $\rightarrow$ Common Law Assize $\rightarrow$ The Ninth Hanging. \emph{Cracks}: produce a witnessed Brehon judgment older than the crown's claim; make the kin-whole with cattle, not counted coin; let the Green Host take his name. \textit{Grey Liturgy variant}: tie a grey thread around the noose. The rope will not tighten. The sheriff will hang alone.
\item \textbf{The Abbess's Golden Inquisition (Everflame purifier)} --- \emph{Phases}: Doctrine of Silence $\rightarrow$ Trial by Censer $\rightarrow$ Brand of Orthodoxy. \emph{Cracks}: expose her pact with raiders; turn her own clerics against her with truth; let the Green Host ride during the trial. \textit{Grey Liturgy variant}: ring a Confessor's bell in the cathedral. The Abbess will hear. She will not confess. The bell will ring anyway.
\item \textbf{The Missing Step (living omission)} --- \emph{Phases}: Vanished Witness $\rightarrow$ Unread Clause $\rightarrow$ Unwalkable Bridge. \emph{Cracks}: invite a Taboo-Witness to count aloud; spill nine drops into Saint's Well but \emph{name only eight}; let the Lady Beneath keep a memory. \textit{Grey Liturgy variant}: step where the step is missing. You will fall. You will land on a grey thread. The Confessor will pull you back. You will not remember where you fell.
\end{itemize}