\section{Thepyrgos --- ``The City of Unfinished Proofs''}
\label{chap:thepyrgos}

\textbf{Key Demographics: } Humans
\textbf{Capital City: } Thepyrgos (The City)

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{Every stair is a question. Every answer is a proof. The Ninth and the Witness argue over whose record holds the truth.}

\begin{quote}
  \textbf{The Archon (Elite):}  
  ``Every stair climbed is a word in the great conversation between earth and sky. This city is not built---it is grown, stone by stone, law by law, until the very architecture speaks wisdom. But the stairs remember more than the living. They remember the ones who never came down.''
\end{quote}

\begin{quote}
  \textbf{Bell-Runner (Commoner):}  
  ``I know every step that creaks a warning and every landing that whispers gossip. The bells tell the city's mood better than any lord---three chimes means business, six means trouble, and nine means the hunters are coming. But on the ninth bell, the stairs themselves answer.''
\end{quote}

\noindent\textbf{Genre / Mood:} Scholarly vertical city, labyrinthine imperial bureaucracy, coastal peninsular folk horror in the hinterlands, metaphysical war between memory and forgetting.

\noindent\textbf{Starting Location:} The Synod Hall at dusk. Gold-glass mosaics glitter as a Matriarch reads a scroll aloud. ``The Beacon Crown sees traitors in the smoke. A child's talent was surrendered to the Commissary last night. Find the informer---or the next stair seized will be yours. And if you hear a stair creak where none exists, do not turn around. The Ninth is counting.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
Thepyrgos is both realm and city, built on the bones of the drowned \textbf{Astroegro Empire}. The capital rises in tiers of towers, terraces, and endless stairs above the Dolmis shore---but the stairs are not stone. They are \emph{fossilized vertebrae}, the spine of a leviathan that Khemesh dragged under when the last emperor broke a covenant older than the sea. Once a Utaran province, now a nation unto itself, Thepyrgos is famed for its universities, libraries, and debating synods, where precedent and philosophy weigh as much as pikes or coin.

The Astroegro were a human civilization of formidable magic, the ancient rivals of the \textbf{Valewood's Empire of Leaves}. Their ruins litter the hills and swamps south of the city, and their legacy haunts every threshold. When the Empire of Leaves fell, \textbf{Lethai-thora} refugees---high elves who had served as archivists and lore-keepers---fled across the Black River and were granted sanctuary in Thepyrgos. Generations later, their descendants still walk the city's tiers, their silver hair and ink-stained fingers a quiet reminder of a pact older than the synod. Many now hold chairs at the university, tend the forbidden archives, or secretly serve the Witness or the Ninth.

Outside the city, the countryside is a rugged coastal landscape of limestone hills, olive groves, and whitewashed villages---but the weather is warmer, wetter, and the valleys are choked with reed-filled marshes. This is the land of the \textbf{Hyro} people, the surviving descendants of the Astroegro. The Hyro carry the fracture of their ancestors in their blood. A child born in the hills might dream in geometries that do not exist. A farmer might hear the stones whispering in a language no living throat can shape. The \textbf{Chain-Lanterns} and other Witch-Hunter Commissariats ruthlessly patrol these villages, using posted warrants, candle-smoke tests, and stair-seizures to identify and kidnap children with arcane talent. The Hyro have learned to offer \emph{substitutes}---a memory, a false name, a lock of hair---at crossroads, hoping the Hollow accepts the tribute.

In the university's forbidden tiers, the \textbf{Witness} and the \textbf{Ninth} fight a proxy war via fraternal orders and occult mystery cults. The Witness wants the \textbf{Astroegro Vaults} \emph{remembered} but sealed; the Ninth wants them \emph{opened}, to trigger a Great Unlearning that would erase the concept of record from the world. No provenance would remain. The Hollow would go unanswered. The Witness would have nothing to inscribe. The Ninth would become the only law: \emph{forgetting}. No one is sure which side the Archon secretly favors. And in the ruins of the Astroegro, something still ticks---gears of black glass, locked doors that hum, and the occasional survivor with eyes like burning tar who speaks of a pressure from below.

\vspace{2mm}
\textbf{What you notice first:} The endless chime of bells from different tiers, never quite in sync. The smell of brine, olive oil, and candle wax. The way locals glance at the sky when a child coughs---and then at the nearest crossroads. The quiet grace of the Lethai-thora archivists, who never seem to blink.

\textbf{The one rule that kills newcomers:} Never let a bell-ringer count your breaths. The Chain-Lanterns offer safe conduct to those who give names that ring false on the test stairs. But worse: never ask a question of a stair that has no end. The Ninth is listening, and it will answer.
\end{tcolorbox}

\subsection*{Regional Mood: Stairs, Bells, and the Argument of Patrons}

Here, every ascent is a plea and every descent a confession. Thepyrgos the city is a vertical labyrinth of stair-rights, rope-guilds, and synod edicts that shift with the tide. Thepyrgos the province is a haunted countryside of hidden talent and state-sanctioned abduction. The Hyro keep their arcane children in root cellars and haylofts, praying the hunters take the neighbor's instead. The Chain-Lanterns move at dawn, their brass censers swinging, their bells tuned to the frequency of latent magic. In the university's forbidden tiers, the Witness and the Ninth arm rival fraternities with citations and curses, and no one is sure which side the Archon secretly favors. The Lethai-thora maintain the great archives, but they are divided---some serve the Witness as faithful recorders, others whisper to the Ninth in darkened scriptoria. And in the ruins of the Astroegro, something still ticks---gears of black glass, locked doors that hum, and the occasional survivor with eyes like burning tar who speaks of a pressure from below.

\subsection*{Geographic Integration}

Thepyrgos is landlocked on three sides by Viterra, separated by the Black River and the Belmont Marshes. The marshy lowlands breed strange fogs and rarer creatures, and the border is porous---Viterran hedge-courts and Thepyrgos synods argue jurisdiction over every flooded farmstead. To the northeast, a short boat ride from Fairport (Viterra) across the Belworth estuary brings travelers to Marcott in Vhasia, making Thepyrgos a quiet but crucial link in the riverine trade between the fractured sun-kingdom and the imperial remnant. The city itself controls the \textbf{Astroegro Straits}---the narrow sea passage between the Dolmis and the Amaranthine---and maintains a few small colonies on the western coast of Hyaar, where olive oil and wine are exchanged for eastern silks and spices. Those colonies are run by exiled scholars and retired Chain-Lanterns, and they are said to be even more paranoid than the mother city.

\subsection*{Faction Timers (Thepyrgos)}
\label{sec:thepyrgos-faction-timers}

The politics of Thepyrgos are a delicate balance between four powers: the Archon's civil authority, the Synod's religious law, the Commissariat's witch-hunters, and the hidden war between the Witness and the Ninth.

\begin{itemize}
\item \textbf{Archon's Balance [6]} --- Tracks the Archon's ability to keep the synod, guilds, and Commissariat in check.\\
  Advances when: the party exposes a Commissariat secret (+1), a Synod edict is defied (+1), or a faction gains too much influence (+1).\\
  At 4 segments: the Archon's authority frays; all social rolls involving civic authority gain +1 DV.\\
  At 6 segments: the Archon loses control of one faction---either the Commissariat runs wild, the Synod calls for his deposition, or the Ninth's Coterie operates openly. The party must choose which faction to support.

\item \textbf{Commissariat Pressure [4]} --- Tracks the Chain-Lanterns' hunting intensity.\\
  Advances when: the party uses unsanctioned magic in the city, helps a gifted child escape, or defies a Commissary warrant (+1).\\
  At 4 segments: the Commissariat conducts a city-wide sweep. All PCs must test Spirit+Resolve (DV 4) or be detained for questioning. A detainee may be released for 1 Boon or a favor to the Commissariat.

\item \textbf{Ninth's Progress [8]} --- Tracks the Ninth's Coterie's efforts to open the Astroegro Vaults.\\
  Advances when: the party opens a vault (+2), speaks a true name of an Astroegro ruin (+1), or accepts a boon from the Ninth (+1).\\
  At 4 segments: the Coterie finds a vault entrance. At 8 segments: they open a vault. The Unfinished Stair appears in the city. The party must decide: help them, stop them, or seal the vault themselves.

\item \textbf{Witness's Vigil [6]} --- Tracks the Witness's Archivists' efforts to record and seal dangerous knowledge.\\
  Advances when: the party records a truth in the Synod's archives (+1), seals a vault (+2), or refuses a boon from the Ninth (+1).\\
  At 4 segments: the Witness's Archivists offer the party a boon (a free Synod Indulgence or Archive Hour). At 6 segments: they demand a favor---record a dangerous truth or face their scrutiny.
\end{itemize}

\subsection*{Faction Relationship Dial: Witness ↔ Ninth}
\label{sec:thepyrgos-dial}

Track the party's perceived allegiance between the Witness (memory, record, preservation) and the Ninth (forgetting, erasure, revelation).

\begin{itemize}
\item \textbf{Witness-leaning:}
  \begin{itemize}
    \item +1 Position on social rolls with archivists, Synod clerks, and the Matriarch
    \item The Chain-Lanterns view you as potential informants (first interrogation starts at +1 Position)
    \item The Ninth's Coterie actively works against you (complications involving forgotten records or hidden stairs)
  \end{itemize}

\item \textbf{Ninth-leaning:}
  \begin{itemize}
    \item +1 Effect when seeking hidden knowledge, forbidden texts, or vaults
    \item The Commissariat views you as dangerous (first contact starts Desperate)
    \item Archivists are wary of you; the Synod may call you to account
  \end{itemize}

\item \textbf{Centering actions:}
  \begin{itemize}
    \item Record a truth in the Synod's archives (move toward Witness)
    \item Feed a memory to the Ninth's Coterie (move toward Ninth)
    \item Perform a ritual that acknowledges both (center the dial, clear 1 segment from Unfinished Stair)
  \end{itemize}
\end{itemize}

\subsection*{Prominent NPCs of Thepyrgos}
\label{sec:thepyrgos-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Archon Konstantinos & Ruler of the city, elected by the wards & Balance the synod, the guilds, and the Commissariat & He wears a key around his neck; no one knows what it opens---or that it opens the Ninth's Vault. \\
\hline
Matriarch Eleni & Synod primate, head of the Temple of Light's synod & Keep the Everflame from reclaiming the province & She speaks once a session; when she does, the bells listen. She has never spoken the ninth word. \\
\hline
Commissary Kyrillos & Chain-Lantern commander & Purge all untutored magic from the province & His lantern shows not only latent talent but also the bearer's greatest fear. He refuses to look into it himself. \\
\hline
Sofia & Bell-Runner (commoner exemplar) & Outrun the hunters and protect the hidden gifted & She ties a knot in her hair for each hunter she has outrun; her hair is a thicket. One knot is white---a hunter she could not outrun. \\
\hline
Nomophylax Zoe & Law-scholar, master of citations & Reform the synod from within & She has never lost a case; her opponents have a habit of vanishing into the Unfinished Stair. \\
\hline
The Lighthouse-Patriarch & Reclusive seer of the Beacon Crown & Steer the city through the coming storm & He has no eyes---only mirrors where his eyes should be. In the mirrors you see the faces of those whose names the Ninth holds. \\
\hline
Keeper Lian (Lethai-thora) & High Archivist of the Library of Keys & Preserve the Astroegro records and keep the Ninth's cult from the deepest vaults & She has not spoken aloud in twenty years. She writes on water. The water never freezes. \\
\hline
\end{longtable}

\subsection*{The Archon's Key}
\label{sec:thepyrgos-archon-key}

The Archon wears a key around his neck. No one knows what it opens. The key is a stair.

\begin{itemize}
\item \textbf{What it opens:} The Ninth Vault---the deepest of the Astroegro vaults, sealed by the first Archon. It contains the original Astroegro orrery, a working model of the cosmos that still ticks.

\item \textbf{What happens if the party uses it:}
  \begin{itemize}
    \item The vault opens. The Ninth's Coterie is immediately aware.
    \item The Witness's Archivists arrive within the hour.
    \item The party must choose: destroy the vault's contents (gain the Witness's favor), use them (gain the Ninth's favor), or seal it again (gain the Archon's favor---and a permanent enemy in whichever faction they defied).
  \end{itemize}

\item \textbf{How the party can learn this:}
  \begin{itemize}
    \item A dying Archivist whispers it (cost: 1 SB)
    \item The Lighthouse-Patriarch shows them in a mirror (cost: 1 Fatigue)
    \item The Ninth's Coterie offers to trade the information (cost: 1 Obligation to the Ninth)
  \end{itemize}
\end{itemize}

\subsection*{The Lighthouse-Patriarch's Mirror}
\label{sec:thepyrgos-patriarch}

Once per session, a character who spends a scene in the Lighthouse-Patriarch's presence may ask the GM one question about:
\begin{itemize}
\item A hidden door or stair in the city
\item The location of an Astroegro vault entrance
\item A weakness of the Chain-Lanterns or the Ninth's Coterie
\end{itemize}

\textbf{Cost:} The mirrors show the asker a face they have wronged. Mark 1 Fatigue (guilt).

\textbf{Limit:} The Patriarch will not answer the same question twice. He does not need to speak---the mirrors show the answer. If the party asks a question the Patriarch cannot answer, the mirrors show a blank expanse of water.

\subsection*{Names of Thepyrgos}
\label{sec:thepyrgos-names}

Inspired by the southern peninsulas and the Aegean coast: clear vowels, soft endings, and surnames that recall a stair, a tower, or a village. Lethai-thora names are longer, more flowing, often ending in -iel, -ael, -ir.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Konstantinos & Eleni & Skalas & \textit{the Stair-Wise} \\
\hline
Dimitris & Zoe & Platanos & \textit{the Unseen} \\
\hline
Kyrillos & Sofia & Raptis & \textit{Bell-Sworn} \\
\hline
Nikolas & Anastasia & Laskaris & \textit{the Accursed} \\
\hline
Leonidas & Irini & Makris & \textit{Ash-Finger} \\
\hline
Stephanos & Fotini & Koutroulis & \textit{the Silent} \\
\hline
Aerendil & Lian & Thoraen & \textit{Archivist} \\
\hline
\end{longtable}

\paragraph*{(Tower/Stair/Sea-wall)} Pilgrim's Stair zig-zagging sea-cliff to wicket gate; Tower Quarter with family keeps stitched by ladders and bridges.

\section*{Spades --- Places}
\label{sec:thepyrgos-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Pilgrim's Stair} --- Zig-zagging sea-cliff to wicket gate; ascent as devotion. `[TRAVEL]` `[RITUAL]`
  \hfill \textit{Every hundredth step has a bell. Pilgrims ring once for mercy, twice for memory. The ninth bell has no clapper.}
\item \textbf{Tower Quarter} --- Family keeps stitched by ladders and bridges; vertical neighborhoods. `[SOCIAL]`
  \hfill \textit{The lowest rung of each ladder is painted red---a mark of the rope-guild's claim. The rope-guild's scroll is written in disappearing ink.}
\item \textbf{Chain-Barbican} --- Harbor gate; capstans ready to raise the boom at horn's call. `[AUTHORITY]`
  \hfill \textit{The capstan pins are carved from the thighbones of executed smugglers. The bones still warm in your hand.}
\item \textbf{Blue Cistern} --- Under the forum; cool echo, strict tally; water as law. `[MAGIC]` `[PROOF]`
  \hfill \textit{The water tastes of copper and old prayers. Drink deep, and you may forget a formula---but the water will remember your face.}
\item \textbf{Ropeyard Terrace} --- Cranes walk cargo up slope by treadwheel; muscle and machinery. `[LABOR]`
  \hfill \textit{The treadwheel operators sing in eights. The ninth beat is always silent---that is when the Stair moves.}
\item \textbf{Storm-Wall Arcades} --- Wind shrieks through arched galleries; defense as architecture. `[WEATHER]`
  \hfill \textit{The wind carries voices from the last siege. Some say they're warnings. Others say they're the Ninth rehearsing.}
\item \textbf{Beacon Crown} --- Ring of watchtowers signaling along the Dolmis; eyes of the city. `[TRAVEL]` `[AUTHORITY]`
  \hfill \textit{The beacons burn juniper and saltpeter. Black smoke means fire; white smoke means hunters. Green smoke means the Stair is hungry.}
\item \textbf{Library of Keys} --- Archive vaults; shelves locked by rung and rite; knowledge as power. `[MAGIC]` `[SECRET]`
  \hfill \textit{The keys are graded by the import of what they lock. The ninth key opens nothing---and everything. It is kept in a box that has no lid.}
\item \textbf{Siege Court} --- Test ramp and scarred stones; war never far from peace. `[COMBAT]`
  \hfill \textit{The ramp's stones are stained with lamp oil and regret. Practice charges end in the cistern---or the Unfinished Stair.}
\item[J] \textbf{Ladder Monastery} --- Perched on a needle of rock; monks climb toward truth. `[RITUAL]`
  \hfill \textit{The monks never speak. They write commandments in chalk on the rungs. The ninth rung has no chalk---only a smear of ash.}
\item[Q] \textbf{Synod Hall} --- Gold-glass mosaics; judgment audible at a whisper's distance. `[LAW]`
  \hfill \textit{The acoustics are enchanted. A lie told here echoes twice---once in the liar's voice, once in truth's. The Ninth's lies echo only once.}
\item[K] \textbf{Archon's Citadel} --- Acropolis peak; mirror tiles glare like noon's justice. `[AUTHORITY]`
  \hfill \textit{The mirrors were a gift from Kahfagia. They show not the sea, but the faces of those who bear the archive's mark---and sometimes a stair that was not there before.}
\item[A] \textbf{The Unfinished Stair} --- A hidden flight that appears only to those who have asked a question the Synod cannot answer. `[THRESHOLD]` `[MYSTERY]`
  \hfill \textit{The stairs have no end. Each step is a different year. At the top, the Ninth and the Witness are playing knucklebones with forgotten names.}
\end{enumerate}

\paragraph*{(Archon/Synod/Guild)} Bell-runner child whose feet know every stair by heart; Master of Ropes with palms like oak, voice like a whistle.

\section*{Hearts --- People \& Factions}
\label{sec:thepyrgos-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Bell-Runner} --- Child whose feet know every stair by heart; speed as currency. `[STEALTH]`
  \hfill \textit{She ties a knot in her hair for each hunter she has outrun; her hair is a thicket. One knot is white---a hunter she could not outrun.}
\item \textbf{Master of Ropes} --- Palms like oak, voice like a whistle; guild head of lifters. `[LABOR]`
  \hfill \textit{He lost two fingers to a frayed line. He keeps them in a jar as a warning. The jar is on the ninth shelf.}
\item \textbf{Icon-Smith} --- Portable shrine, temper about pigment; faith in detail. `[CRAFT]`
  \hfill \textit{His pigments are ground from stolen icons. He calls it recycling the divine. The Ninth calls it preparation.}
\item \textbf{Wall Strategos} --- Drills levies on the test ramp; defense as daily ritual. `[COMBAT]`
  \hfill \textit{She lost a son to a Linn raid. She trains orphans now---and keeps a list. The list has nine names.}
\item \textbf{Chain-Keeper} --- Harbor boom; hates ``surprises'' at dusk; order as obsession. `[AUTHORITY]`
  \hfill \textit{His bell never rings the same note twice. He says the tide changes its mind. The Ninth changes its tone.}
\item \textbf{Oath-Examiner} --- Weighs words like iron; notary of truth and consequence. `[LAW]`
  \hfill \textit{She carries a scale with one pan empty. The empty pan is for lies. When the Ninth speaks, both pans tip.}
\item \textbf{Archive Sister (Lethai-thora)} --- Key to the ``forbidden rung''; keeper of dangerous knowledge. `[SECRET]`
  \hfill \textit{Her left eye is milk-white---a cost for reading a cursed shelf. The shelf was a stair in a past life.}
\item \textbf{Salt-Fish Syndic} --- Warehouse queen of the quays; profit in preservation. `[TRADE]`
  \hfill \textit{Her salt is cut with ground bone meal. The fish stay fresh---and quiet. The bones were from a stair.}
\item \textbf{Nomophylax} --- Law-scholar whose citations close mouths; precedent as weapon. `[LAW]`
  \hfill \textit{He has never lost a case. His opponents have a habit of vanishing. The Unfinished Stair has a new step.}
\item[J] \textbf{Palikar Captain} --- Tower guards; ladder-quick and letter-shy; strength in motion. `[COMBAT]`
  \hfill \textit{His shield is painted with a ladder---each rung a battle he survived. The ninth rung is blank.}
\item[Q] \textbf{Matriarch of the Ladder} --- Synod primate; calm as stone, wisdom in height. `[AUTHORITY]`
  \hfill \textit{She speaks once a session. When she does, the bells listen. She has never spoken the ninth word.}
\item[K] \textbf{The Archon} --- Elected, entrenched, counting bells; ruler of the ascent. `[AUTHORITY]`
  \hfill \textit{He wears a key around his neck. No one knows what it opens. The key is a stair.}
\item[A] \textbf{Lighthouse-Patriarch} --- Reclusive seer; said to steer beacons in storms. `[MYSTERY]`
  \hfill \textit{He has no eyes---only mirrors where his eyes should be. In the mirrors you see the Unfinished Stair.}
\end{enumerate}

\paragraph*{(Edict/Quake/Chain)} Tremor with hairline cracks racing down a stair; crowds freeze; iconoclast riot in the Tower Quarter.

\section*{Clubs --- Complications/Threats}
\label{sec:thepyrgos-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Tremor} --- Hairline cracks race down a stair; crowds freeze, dust speaks. `[DISASTER]`
  \hfill \textit{The dust spells names. Yours appears if you've broken a stair-right---or if the Ninth is curious.}
\item \textbf{Iconoclast Riot} --- In the Tower Quarter; pigments and teeth fly through the air. `[SOCIAL]` `[COMBAT]`
  \hfill \textit{The riot started over a missing icon's eye. The eye was a hunter's informer. The icon was a stair.}
\item \textbf{Chain Jam} --- At the barbican; capstan pins shear at the worst moment. `[TRAVEL]`
  \hfill \textit{The pins were filed by a saboteur. The chain will hold---or not. The Ninth is keeping score.}
\item \textbf{Black Northerly} --- Pins cranes; cargo hangs between tiers, prayers below. `[WEATHER]`
  \hfill \textit{The wind smells of iron and ice. Someone is using weather magic. The Witness disapproves.}
\item \textbf{Cistern Taint} --- Sour water; inspectors padlock spouts, thirst rises. `[PROOF]`
  \hfill \textit{The taint is sweet---too sweet. Someone poisoned the cistern to flush out a talent. The talent is a stair.}
\item \textbf{Synod Summons} --- A chant or badge is suddenly proscribed; law shifts like wind. `[LAW]`
  \hfill \textit{The summons is written on a bell-clapper. Read it, and you are a witness. The Ninth writes the margin.}
\item \textbf{Rope Guild Interdict} --- No lifts until a grievance is heard; the city stops climbing. `[LABOR]`
  \hfill \textit{The guild's grievance is a single name. The name is yours. The name is also a stair.}
\item \textbf{Siege Drill Panic} --- Practice horns mistaken for real; chaos in the ranks. `[FEAR]`
  \hfill \textit{The horns were blown by a child who didn't know the difference. Or did they? The child is a step.}
\item \textbf{Smuggler's Ladder} --- Discovered; the whole stair sealed, routes rerouted. `[STEALTH]` `[PROOF]`
  \hfill \textit{The ladder was built by a Hyro family. Now they are all ``guests'' of the Commissariat. Their house is a stair.}
\item[J] \textbf{Ropeyard Fire} --- Sparks leap uphill faster than boots; smoke chokes the ascent. `[DISASTER]`
  \hfill \textit{The fire was set to cover a kidnapping. The victim's name is on the proscription list. The board is a stair.}
\item[Q] \textbf{Exarch's Claim} --- ``Thepyrgos under my protection''; arrests follow the papers. `[AUTHORITY]`
  \hfill \textit{The Exarch is a puppet of the Chain-Lanterns. His seal is a bell without a clapper. The Ninth has the clapper.}
\item[K] \textbf{General Watch} --- All towers manned; levies seize carts ``for the walls.'' `[AUTHORITY]`
  \hfill \textit{The levies are looking for one thing: children who flinch at candle-smoke. The children are stairs.}
\item[A] \textbf{The Unfinished Stair Awakens} --- A hidden flight appears in the middle of a crowded tier. The Ninth's attention falls. `[THRESHOLD]` `[BOSS]`
  \hfill \textit{The stairs have no end. The Witness is writing. The Ninth is counting. You are already on the third step.}
\end{enumerate}

\paragraph*{(Key/Rite/Paper)} Stair token passes a barricaded stair without delay; harbor pass lowers the boom for your vessel on command.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:thepyrgos-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Stair Token} --- Pass a barricaded stair without delay (once); height as right. `[TRAVEL]`
  \hfill \textit{The token is a bronze step. It fits into a missing rung. The missing rung is a secret.}
\item \textbf{Harbor Pass} --- Boom lowers for your vessel on command; sea as ally. `[TRAVEL]`
  \hfill \textit{The pass is a note written on a fish scale. The gate-keeper will not ask why. The fish was a witness.}
\item \textbf{Cistern Draw-Right} --- A named ward during scarcity; water as favor. `[SOCIAL]`
  \hfill \textit{The right is a clay cup. Break it to claim your share. The cup was a stair.}
\item \textbf{Crane Allotment} --- One guaranteed lift on the Ropeyard Terrace; height as service. `[LABOR]`
  \hfill \textit{The allotment is a knotted rope. Cut a knot for each use. The ninth knot is a lie.}
\item \textbf{Icon License} --- Display a disputed image without penalty; faith as leverage. `[SOCIAL]`
  \hfill \textit{The license is stamped with the Archon's ring. The ring is warm---the Archon is watching. The ring is a stair.}
\item \textbf{Archive Hour} --- Library of Keys; one shelf opened to you; knowledge as power. `[SECRET]`
  \hfill \textit{The shelf is sealed with a riddle. Solve it, or the key will break. The riddle is about the ninth step.}
\item \textbf{Bellmark} --- Lawful ring at a fog-wick summons the watchers; law travels on sound. `[AUTHORITY]`
  \hfill \textit{The bellmark is a brand on your palm. Show it, and the watch bleeds for you---once. The blood is a witness.}
\item \textbf{Watchlight Code} --- Today's beacon and mirror signals; navigation as paper. `[TRAVEL]`
  \hfill \textit{The code is written in lemon juice. Heat it to read, then hide it again. The heat is a confession.}
\item \textbf{Synod Indulgence} --- One proscribed chant or rite permitted (for you). `[MAGIC]`
  \hfill \textit{The indulgence is a bell-note held in a jar. Break the jar to release it. The bell-note is a name.}
\item[J] \textbf{Pronoia Grant} --- Temporary stewardship of a terrace hamlet; land as standing. `[TRADE]`
  \hfill \textit{The grant is a wooden token. Present it to the village elder---if she is still alive. The elder is a stair.}
\item[Q] \textbf{Synod Audience} --- Private with the Matriarch and clerks; whispers become policy. `[SOCIAL]`
  \hfill \textit{The audience is held in a fog hollow. What is spoken there is never written. The fog is a witness.}
\item[K] \textbf{Archon's Writ} --- Palikars escort you between towers; authority as escort. `[AUTHORITY]`
  \hfill \textit{The writ is a bell-rope. Tie it to your belt, and the guards will know your voice. The rope was a stair.}
\item[A] \textbf{Golden Key} --- Open or seal any city gate once; access as power. `[MAGIC]`
  \hfill \textit{The key is a finger bone. The lock is a gauntleted fist. It will work---once. The fist is the Ninth's.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:thepyrgos-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (keys/rites/papers) that change position rather than call for a roll.
\item If any A appears, echo \textbf{height and sound} motifs---bells that carry, wind that remembers, stairs that echo with purpose.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:thepyrgos-mechanics}

\begin{itemize}
\item \textbf{Bell-Law:} Bells are law made sound. Once per session, a Bellmark or tower bell summons the watch, a levy, or a decree. False rings stain a reputation; the Synod may inquire. The ninth ring is never false---it is always the Ninth.

\item \textbf{Stair-Rights:} Every stair is claimed, chartered, or under review. A token bypasses barricades and checkpoints; climbing the wrong stair without leave sparks a riot or a duel. The Unfinished Stair has no owner---only one who bears its mark.

\item \textbf{The Unfinished Stair Timer [8]:} Tracks the party's entanglement with the hidden architecture of Thepyrgos.
  \begin{itemize}
    \item 0-2: The stairs are hidden. The party has not yet attracted the Ninth's attention.
    \item 3-4: The stairs begin to appear. At night, the party sees an extra step where none should be. The Ninth's Coterie begins to take notice.
    \item 5-6: The stairs appear in daylight. A character who walks an Unfinished Stair may ask one question of the Ninth---but the Ninth will ask one in return. The party is marked. The Chain-Lanterns receive visions of their faces.
    \item 7-8: A vault entrance opens. The Ninth's Coterie is already there. The party must choose: enter the vault, seal it, or destroy it. The Unfinished Stair is now visible to all.
  \end{itemize}
  Tick the timer when you use a hidden stair, ask a question the Synod cannot answer, or speak a true name of an Astroegro ruin. Reduce the timer by feeding a memory to the Ninth's Coterie (lose a fact, a skill point, or a Bond) or by recording a truth in the Synod's archives (gain the Witness's Annotation).

\item \textbf{The Offering at the Crossroads (Hedge Gift):} Once per session, a Hyro character (or someone who has learned their customs) may leave a \emph{substitute} at a crossroads---a written name, a lock of hair, a childhood memory. For the next scene, they are \emph{invisible to the Chain-Lanterns} (cannot be tracked or identified by candle-smoke). The substitute becomes an \emph{Unfinished Proof} that the GM may call in later (the Hollow accepted it; the Hollow will ask for more).

\item \textbf{The Candle-Smoke Test (GM Tool):} When the Chain-Lanterns test a village, the GM rolls 1d6. On a 1-2, they find a gifted child (the party must intervene or witness a tragedy). On a 3-4, they find a substitute (a false name, a hidden memory) and leave. On a 5-6, the village has already paid in silence---no raid occurs. The party can influence the roll by offering their own substitute (use a Boon to shift the result by 1).

\item \textbf{The Astroegro Singing (Player Tool):} A Hyro character can hum a fragment of an old empire song. The song has no words---only pressure. When hummed at a ruin, it may open a hidden door. When hummed at a Chain-Lantern, it may cause their lantern to flicker. Each use advances the \emph{Unfinished Stair} timer by 1 (the past is waking).
  \begin{itemize}
    \item \textbf{Consequences (d6):} Roll each time the song is used.
    \item 1: The song attracts a Chain-Lantern patrol. They arrive in 1d4 scenes.
    \item 2: A stone cracks nearby. A hidden door opens---or closes.
    \item 3: A ghost of an Astroegro archivist appears. It will not speak, but it will point.
    \item 4: The singer forgets the song. They cannot use it again until they learn a new fragment.
    \item 5: The Hollow responds. The singer gains the Condition \textit{Heard by the Hollow} (the GM may spend 1 SB to have a Oath-Keeper appear).
    \item 6: The song is heard by the Ninth's Coterie. They will offer a bargain within the day.
  \end{itemize}

\item \textbf{Lethai-thora Archivists:} The high elf refugees have integrated into the city's scholarly elite. They are the keepers of the deepest archives and the most secret keys. A character with Lethai-thora heritage (or who earns their trust) may gain access to the \emph{Archivists' Mark}---a brass key that opens one restricted door in the Library of Keys per session. However, the Archivists are not monolithic; some serve the Witness, others the Ninth. Choosing a side may alienate the other faction.

\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Unwritten Citation} --- The Witness. \textit{A precedent that exists only in the margin.}
    \item \textbf{The Bell-Founder} --- Mykkiel. \textit{The law that rings true, even when it hurts.}
    \item \textbf{The Abyssal Pressure} --- Khemesh. \textit{The pressure beneath the city; the pressure that the deep remembers.}
    \item \textbf{The Silent Stair} --- The Ninth. \textit{The step that is never counted, the threshold that opens only once.}
  \end{itemize}
\end{itemize}

\subsection*{The Astroegro Vaults --- Mini-Generators}
\label{sec:thepyrgos-vaults}

Beneath the city, nine vaults. The Synod knows of three and has sealed them with Witness-branded wax. The Ninth's Coterie has found a fourth. The other five are still buried, waiting for the right key---a true name, a specific death, a memory that has never been spoken.

\subsubsection*{Vault Contents (d6)}
When the party opens a vault (or finds a way in), roll to determine what lies within.

\begin{enumerate}
\item \textbf{The Echo Forge} --- A chamber of black glass gears that still turn. The forge produces a single perfect bell every nine years. The current bell is half-formed and rings at a frequency that makes teeth ache. \emph{Weapon:} A bell that, when rung, forces all who hear it to speak one true name (Resolve DV 4 to resist). \emph{Guardian:} A ghost-smith who does not attack---it asks you to complete the bell. \emph{Cost:} A year of your life to finish the forging.

\item \textbf{The Unwritten Archive} --- Shelves of blank stone tablets. The tablets are not blank---they are written in a script that only the dead can read. A ghost-archivist sits at a desk, writing the same number over and over. \emph{Weapon:} A quill that writes in disappearing ink; the words become true for one hour, then vanish. \emph{Guardian:} The archivist's ghost will answer one question about the vault's history, but will then ask for a number you have never written. \emph{Cost:} You must forget how to read for one day.

\item \textbf{The Glass Orrery} --- A working model of the Astroegro cosmos, its gears carved from abyssal glass. It still ticks. The ticking changes when you lie. \emph{Weapon:} A lens that shows the death of anyone you name---but only once per person. \emph{Guardian:} A pressure-spirit that coils around the orrery; it will let you take the lens if you can answer its riddle. \emph{Cost:} A dream of drowning (you will dream of the sea every night until you return).

\item \textbf{The Bone Gallery} --- A corridor lined with the remains of Astroegro archivists. Each skeleton wears a different style of bell-maker's apron. The newest apron is still damp. \emph{Weapon:} A finger bone that, when pointed at a door, opens it---once. \emph{Guardian:} The archivist whose bone you take will ask for a story about a door you should not have opened. \emph{Cost:} The archivist's ghost follows you until you close a door you left open.

\item \textbf{The Counting Chamber} --- A room where the walls are covered in tally marks---eight scratches, a gap, eight scratches. The marks are not in sequence. A wright's skeleton still holds a hammer. \emph{Weapon:} A hammer that can drive a copper nail into any surface, sealing a threshold for one scene. \emph{Guardian:} The wright will not attack---it will only count. If you interrupt the count, it becomes hostile. \emph{Cost:} You must leave a tally of your own---a memory, a name, a debt---on the wall.

\item \textbf{The Sealed Heart} --- A chamber of black glass and fossilized bone, where the heat of the earth meets the cold of the void. Something coils in the dark---a child of Nidhoggr, a fragment of the Deep Drake's attention. \emph{Weapon:} A scale that, when touched, grants +2 dice to resist fear for one scene---but the scale grows warm each time. \emph{Guardian:} The pressure that thinks. It does not attack; it invites you to stay. \emph{Cost:} A memory of sunlight. You will forget what warmth feels like for one day.
\end{enumerate}

\subsubsection*{Vault Guardian Quirks (d6)}
In addition to their primary nature, each vault's guardian has a quirk that defines how they can be negotiated with.

\begin{enumerate}
\item \textbf{Counting Obsession:} The guardian will not attack unless you interrupt its count. It counts in eights. If you join the count correctly, it will give you a boon. If you miscount, it takes a memory.
\item \textbf{The Witness's Seal:} The guardian is bound by a Witness-branded wax seal. If you present a truth the Seal cannot refute, the guardian must let you pass. The truth must be witnessed by another living soul.
\item \textbf{Ninth's Silence:} The guardian only speaks in gaps---silences between words, spaces between steps. If you can answer a question in the silence, it will let you pass. If you speak over the silence, it attacks.
\item \textbf{Pressure Echo:} The guardian is a pressure-spirit; it only attacks when you are afraid. If you can pass a Resolve test (DV 4) and show no fear, it will let you pass. On a failure, it manifests and attacks (Cap 3).
\item \textbf{Archivist's Grief:} The guardian is the ghost of an archivist who failed to save a vault. It will let you pass if you can prove you have failed something important---a promise, a mission, a person. It will judge your failure. If it deems it worthy, you pass.
\item \textbf{The Hollow's Claim:} The guardian is Hollow-touched. It does not attack; it offers a bargain: one memory from each intruder, and the vault opens. Refusal means it takes a name instead.
\end{enumerate}

\subsubsection*{Vault Costs (d6)}
Every vault demands a cost to open or to take its treasures.

\begin{enumerate}
\item \textbf{Memory of Salt Water:} You must forget a memory of the sea---a voyage, a ship, a drowning. You will remember that you forgot, but not what it was.
\item \textbf{Dream of Drowning:} You will dream of drowning every night until you return to the vault. Each morning, mark 1 Fatigue.
\item \textbf{Confession of Hubris:} You must speak aloud a time you were wrong and knew it, but refused to admit it. The confession must be witnessed.
\item \textbf{A True Name:} You must give the vault a true name---yours, or someone else's. The name will be added to the vault's records.
\item \textbf{A Year of Life:} You will age one year instantly. Gain 1 permanent Condition (e.g., gray hair, a limp, a tremor).
\item \textbf{A Bond Broken:} You must sever a Bond---a relationship, a promise, a debt. The loss is permanent, though the wound may heal.
\end{enumerate}

\subsection*{The Province --- Mini-Generators}
\label{sec:thepyrgos-province}

Beyond the city walls, the hinterlands of Thepyrgos are a haunted countryside of limestone hills, olive groves, and the Hyro people who still remember the Astroegro Empire. The Chain-Lanterns patrol the villages, and the Ninth's Coterie moves through the ruins.

\subsubsection*{Village Situation (d6)}
When the party enters a Hyro village, roll to determine its current state.

\begin{enumerate}
\item \textbf{Quieted Village:} The Chain-Lanterns came last week. The gifted children are gone. The village is silent, and the elders will not speak to outsiders. A single child hides in the root cellar---too old to be taken, too young to understand why.
\item \textbf{Candle-Smoke Test Tomorrow:} The hunters are coming at dawn. The village is in a panic. One family has a child who flinched at the last test. They need a substitute---a lock of hair, a false name, a memory to offer at the crossroads.
\item \textbf{The Ninth's Coterie Visits:} A gray-robed figure has been seen in the village. They offer protection from the hunters in exchange for stories---true names, forgotten memories, fragments of the Astroegro past. The village is divided: some trust them, others fear them.
\item \textbf{The Hollow's Offering:} A child was born last night with eyes the color of the sea. The elders say the Hollow has accepted a substitute---a memory of the Astroegro fall. The child will never have talent. The village is grateful, but the child's mother is weeping.
\item \textbf{Ruin Singing:} The stones outside the village have started humming at night. A Hyro elder believes an Astroegro vault is surfacing---the pressure below is rising. She needs help to listen, to count the stones, to find the door before the hunters hear the song.
\item \textbf{The Bear Cult:} A group of villagers has taken to the hills, wearing cured skins and speaking in growls. They claim the only way to escape the hunters is to become the hunter. They are hunting Chain-Lantern patrols---and they are winning. The Archon has declared them outlaws, but the villagers leave offerings at their shrines.
\end{enumerate}

\subsubsection*{Hyro Encounter (d6)}
When meeting a Hyro villager, traveler, or hermit, roll for their disposition and need.

\begin{enumerate}
\item \textbf{Grieving Elder:} Their child was taken in the last sweep. They will help the party if the party helps them find the child. They know of a hidden stair that leads to the Beacon Crown's secret basement.
\item \textbf{Stone-Listener:} A hermit who hears the pressure below. They can identify a thin place, a vault entrance, or a route that the hunters do not watch. They demand a story about a dream in return.
\item \textbf{Substitute Keeper:} This villager sells false names and stolen memories to families who need to protect their children. They can provide a substitute for a candle-smoke test---but the Hollow will remember the debt.
\item \textbf{Chain-Lantern Informant:} A villager who reports gifted children to the hunters in exchange for protection. They are despised by the village but have useful information about the hunters' schedules. They are also deeply ashamed.
\item \textbf{Ruin Seeker:} A young Hyro who believes the Astroegro vaults contain the key to defeating the hunters. They are reckless, know a few fragments of the old songs, and will follow the party into any ruin.
\item \textbf{Hollow-Touched:} A villager who has given too many substitutes to the Hollow. They have begun to forget their own name. They speak in riddles and can see the Hollow's attention on others. They are useful, but dangerous to be around.
\end{enumerate}

\subsubsection*{Hunter Encounter (d6)}
When the Chain-Lanterns or other Commissariat forces appear, roll for their objective and mood.

\begin{enumerate}
\item \textbf{Routine Sweep:} The hunters are testing children with candle-smoke. They are bored, efficient, and will accept a bribe (1 Boon or 1 Supply) to skip a house. On a Miss, they notice the party's arcane talent.
\item \textbf{Targeted Raid:} The hunters have a warrant for a specific child. They will not be bribed. They will search every house. The party must hide the child, substitute a false name, or fight.
\item \textbf{Inquisitorial Audit:} A Chain-Lantern Captain is auditing the local Commissary. They are looking for corruption, not talent. They may accept information about a corrupt officer in exchange for ignoring a minor talent.
\item \textbf{Ninth Pursuit:} The hunters are not hunting talent---they are hunting the Ninth's Coterie. A gray-robed figure has been seen in the village, and the Chain-Lanterns want to question them. They will detain anyone who looks suspicious.
\item \textbf{Desperate Hunt:} The hunters have lost a prisoner---a gifted child who escaped. They are searching the village house by house. The child is hiding in a root cellar, and the hunters are getting closer.
\item \textbf{Pardon Offer:} The Exarch has offered a pardon to any family that surrenders their gifted child voluntarily. The village is divided---some families are considering it. The party may need to intervene.
\end{enumerate}

\subsubsection*{Hyro Discovery (d6)}
When exploring a ruin, a village, or the hills, roll for something the Hyro have hidden or preserved.

\begin{enumerate}
\item \textbf{Fragment of the Old Song:} A piece of the Astroegro empire's music. Humming it can open one hidden door or cause a Chain-Lantern's lantern to flicker. One use only.
\item \textbf{Substitute Cache:} A hidden stash of false names, locks of hair, and written memories. Enough substitutes for three candle-smoke tests. Using them will not anger the Hollow---this time.
\item \textbf{Kurgan Offering:} A stone marked with a Hyro offering. If you leave a memory here, the Hollow will accept it as a substitute for one child. The memory must be true and painful.
\item \textbf{Warden's Log:} A record of Chain-Lantern patrols, including schedules, informants, and the names of children taken. Valuable to the resistance---and dangerous to carry.
\item \textbf{Sealed Door:} A hidden door in a hillside, marked with an Astroegro seal. Opening it requires a true name---or a memory of the empire. The Hyro have kept it sealed for generations.
\item \textbf{The Singing Stones:} A circle of stones that hum at dusk. They are an Anchor site---a thin place where the pressure below is close. The Hyro leave offerings here. A character who offers a memory may ask one question of the Hollow.
\end{enumerate}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Astroegro Remnant (Khemesh):} Deep beneath the Beacon Crown, a sealed chamber contains a working Astroegro orrery. Its gears are carved from abyssal glass, and it still ticks. Those who listen too long hear a voice whispering the true names of drowned emperors. The Chain-Lanterns have posted a perpetual guard; they say the orrery is a beacon for something that should not surface.
  \item \textbf{Saint Methodia of the Silent Stair (The Witness):} A legendary archivist who climbed the Ninth Stair of the Library of Keys and never came down. Her ghost is said to appear to those who speak a truth the Synod has buried, offering a single citation that can overturn any edict. The cost is a secret of equal measure, and the ghost never forgets.
  \item \textbf{The Cog-Cults of the Clockwork Monad:} Underground fraternities of gear-smiths and logicians who believe the Astroegro Empire fell because its geometry was imperfect. They meet in the basements of the Ropeyard Terrace, trying to reconstruct the lost orrery. The Chain-Lanterns tolerate them because their calculations sometimes predict talent blooms. Heretics from the Everflame call them timer-worshippers; the Cog-Cults call themselves the True Iterators.
  \item \textbf{The Daughters of the Cord (The Bitter Root):} A hospice in the lower tiers is run by gray-robed women who ask no payment, but they take names. They are tolerated because they treat the poor for free, but the Commissariat suspects they also hide gifted children. A renegade Sister hides among the bell-runners, trying to unbind her Hollowed from the Chain-Lanterns' record.
  \item \textbf{The Savage Heart (Red Maw):} In the highland villages west of the city, a bear cult has arisen. They believe the only way to escape the Commissariat's hunts is to become the hunter. They wear cured skins and speak in growls, and they have begun hunting Chain-Lantern patrols. The Archon has declared them outlaws, but the villagers leave offerings of raw meat at their shrines.
  \item \textbf{Zakov's Astroegro Foundation:} The pirate city of Zakov is built on top of a sunken Astroegro port. Smugglers occasionally pull up black glass artifacts from the deep channels, which find their way to Thepyrgos' black market. The Ninth's Coterie offers high standing for such relics; the Witness's archivists try to secure them first.
  \item \textbf{The Ninth:} In any Thepyrgos scene, the ninth bell rung in an hour opens a threshold. The wise stop at eight and cover their ears. Those who ring the ninth are said to see a pale figure standing on the Beacon Crown's mirror tiles---waiting to receive a name.
\end{itemize}

\subsection*{Thepyrgos Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Stair:} Toss a crust onto a rung before climbing. Ignore the first \emph{Tremor} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Rope:} Tie a single knot in a crane line. Cancel \emph{Rope Guild Interdict} or take --1 die on your next lift (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Smoke:} Whisper a hunter's name into the Beacon Crown's fire. Gain +1 Effect when evading the Chain-Lanterns this scene, but gain the Synod's Notice. `[SUPERSTITION]`
  \item \textbf{Offer the Substitute:} Leave a written false name at a crossroads. The next Candle-Smoke Test against you automatically fails---but the Hollow gains an Unfinished Proof on you. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:thepyrgos-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{tower guards or guild masters}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{bell-runners or civic patrols}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{trackers or archive keepers}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable ascent}.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Wall Strategos or guild head} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{quake damage or stair closure}.
\item \textbf{Foe Appears:} A \textbf{rival faction or civic enforcer} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{palikars, guild enforcers, synod guards, or crown militia} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes---\textbf{taboo broken/green host rides/feud declared/crown warrant issued}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{King, Matron, Lady Beneath, or Viterran Governor} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically---\textbf{fog rises/bells ring/causeway shifts/crown forces arrive}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice---\textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Thepyrgos (Height \& Law):} Stairs shift position, bells ring with the wrong tone, legal papers rewrite themselves.
\item \textbf{Thepyrgos (Synod Power):} Edicts change mid-speech, citations appear in the air, wisdom comes unbidden.
\item \textbf{Thepyrgos (Vertical City):} Ropes fray without warning, cranes swing cargo unexpectedly, tower windows show wrong views.
\item \textbf{Thepyrgos (The Unfinished Stair):} A hidden stair appears behind the party. The Ninth is counting. The Witness is writing. The stairs have no end.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth):} On a hit, shift one band. On a strong hit, also impose \emph{Fog Bank}.
\item \textbf{Cut the Stair (Presence + Lore):} On a hit, freeze range for one exchange; if a stair token is in your possession, +1 die.
\item \textbf{Weather the Quake (Body + Athletics):} On a hit in \emph{Tremor}, you pick who keeps their footing; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Bell \& Inquisition)}
\begin{itemize}
\item \textbf{The Bell-Exarch (Nine Bans)} --- \emph{Phases}: Candle-Smoke Sweep $\rightarrow$ Orchestrated Silence $\rightarrow$ Synod Coup. \emph{Cracks}: ring a forgotten bell, expose a false bounty, survive a night in the Beacon Crown.
\item \textbf{The Astroegro Engine (Pressure from Below)} --- \emph{Phases}: Gears Turn $\rightarrow$ Whispers Become Names $\rightarrow$ Seiche. \emph{Cracks}: abyssal glass shard, survivor testimony, the Lighthouse-Patriarch's mirror turning clear.
\item \textbf{The Ninth's Advocate (Living Stair)} --- \emph{Phases}: Count $\rightarrow$ Omit $\rightarrow$ Forget. \emph{Cracks}: speak a true name that the Ninth does not know, offer a memory the Witness has not recorded, climb the Unfinished Stair and refuse to answer the Ninth's question.
\end{itemize}

\subsection*{The Magic Colleges (University Extensions)}
\label{sec:thepyrgos-colleges}
\begin{itemize}
\item \textbf{College of Ladders \& Lines} --- geomancy, architecture, constraint; stairs as sigils, capstans as circles.
\item \textbf{College of Bells \& Beacons} --- acoustics, omen-fire, signal calculus; law made audible, augury made measurable.
\item \textbf{College of Waters \& Weights} --- alchemy, cistern rites, dyes; solutions, precipitates, and lawful flow.
\item \textbf{College of Names \& Oaths} --- nomikon theurgy; citations that bind, clauses that compel.
\item \textbf{College of Mirrors \& Haze} --- illusion, scrutiny; seeing, being seen, un-being seen.
\item \textbf{College of Quiet} --- ethics of death, memory rites; permitted remembrances---anything more is heresy.
\end{itemize}

\subsection*{The Witch-Hunter Province (Commissariat Extensions)}
\label{sec:thepyrgos-hunters}
Thepyrgos-the-Province funds Witch-Hunter Commissariats---ministries with notices posted per ward. Citations hang on stair-landings; bells announce \emph{hunts lawful and hunts less so}. The Chain-Lantern order is the most feared, but others compete for quotas.

\paragraph{Hunter Hearts (faces of the persecution)}
\begin{itemize}
\item \textbf{Commissary-Exarch} --- Writs in triplicate; quotas in private.
\item \textbf{Index-Reader} --- Calibrates children by the bell; ``verification'' by smoke.
\item \textbf{Pardoner-Chaplain} --- Sells absolution to informers; keeps a careful record.
\item \textbf{Stair-Runner Informant} --- Knows which doors open to the right knock.
\item \textbf{Chain-Lantern Captain} --- Commands the most feared order; his lantern reveals truth through flesh.
\end{itemize}

\paragraph{Hunter Clubs (pressures and sweeps)}
\begin{itemize}
\item \textbf{Notice Posting} --- New warrant on ``untutored talent''; crowds turn into nets.
\item \textbf{Candle-Smoke Test} --- Lines at the Blue Cistern; smoke curls toward the Gifted.
\item \textbf{Commissary Sweep} --- Ladders chained; a whole tier searched under horn.
\item \textbf{Quieted Child} --- An ``accident'' near the Exam Stair; evidence goes missing.
\item \textbf{Chain-Lantern Raid} --- Dawn assault on a hamlet; children taken before the rooster crows.
\end{itemize}

\paragraph{Hunter Diamonds (leverage and reprieve)}
\begin{itemize}
\item \textbf{Witness Pardon} --- Informer absolved and commended; cancels one legal fallout \emph{once}.
\item \textbf{Archon's Clemency} --- Suspends a bounty timer for one named; bell-marked and public.
\item \textbf{Hidden Stair Token} --- Through a maintenance run unseen; good for one group escape.
\end{itemize}

\subsection*{Detection \& Denial (Fast Procedures)}
\label{sec:thepyrgos-detect}
\paragraph{Scry \& Deny (evading the tests)}
\begin{enumerate}
\item \textbf{Mask the Trace:} dye, salt, bell-oil; pick \emph{one} (each grants +1 die to one approach).
\item \textbf{Pick the Test:} \emph{Smoke, Sound, Mirror}. State how you spoof it (alchemical veil, bell overtones, angle games).
\item \textbf{On a hit:} cancel a \textbf{Candle-Smoke Test} or \textbf{Proctors' Gate} tag for this scene. \textbf{Consequence:} start \emph{Bell Mark} [2] or \emph{Rector's Notice} [2].
\end{enumerate}

\paragraph{Examination Economy (quick table, GM picks or 1d6)}
\begin{enumerate}
\item \emph{Whisperer} (faint talent): 1 notation; informers abundant.
\item \emph{Spark-Hand} (cantrip): 2 notations; ``rehabilitation'' promised.
\item \emph{Mirror-Bent} (glamour): 3 notations; seized at once.
\item \emph{Bell-Singer} (acoustic): 4 notations; Synod claims jurisdiction.
\item \emph{Oath-Breaker} (binding): 5 notations; Exarch insists on custody.
\item \emph{Quiet-Worker} (death-ethic): 6 notations; College of Quiet petitions for trial.
\end{enumerate}

\subsection*{Scenario Starters (Plug-and-Play)}
\paragraph{The Bright Child on the Exam Stair}
\emph{Hook:} A Bell-Runner begs you to get a latent child to campus before the dawn sweep.  
\emph{Complications:} \textbf{Candle-Smoke Test}, \textbf{Commissary Sweep}, a jealous \textbf{Street Thaum}.  
\emph{Relief:} Trade a \emph{Latency Seal} for testimony; or ring a lawful \emph{Bellmark} to create a corridor---accept \emph{Censure Murmurs} [2].

\paragraph{The Black Index of Accusations}
\emph{Hook:} The College of Quiet wants proof that a Commissary inflates rolls to fill quotas.  
\emph{Complications:} \textbf{Synod Summons}, \textbf{Smuggler's Ladder}, \textbf{Chain Jam}.  
\emph{Relief:} Publish via a \emph{Nomikon} colloquy (gain \emph{Synod Indulgence}); or submit it to the Archon (gain \emph{Archon's Clemency} but start \emph{Faculty Feud} [4]).

\paragraph{The Golden Key and the Silent Stair}
\emph{Hook:} A dying Archivist gives you a Golden Key. ``It opens the stair the hunters fear.''  
\emph{Truth:} The stair leads to the Beacon Crown's secret basement---a prison for children who passed the smoke test.  
\emph{Relief:} Free the children and expose the Commissariat (gain \emph{Archon's Clemency} for the whole party, but start \emph{Exarch's Claim} at full); or seal the stair forever (gain \emph{Golden Key} to use elsewhere, but the children are lost to memory).

\paragraph{The Ninth Vault}
\emph{Hook:} A fragment of an Astroegro weapon has been unearthed in a hill village. The Ninth's Coterie and the Synod both want it. The village elder offers it to the party---if they can offer a substitute that satisfies the Hollow.  
\emph{Truth:} The weapon is a bell that rings the true name of anyone you name. Its guardian is the ghost of the Astroegro archivist who sealed it. The ghost will not speak to anyone who has not climbed the Unfinished Stair.  
\emph{Relief:} Climb the stair, speak the archivist's true name, and take the bell---but the stair adds a segment to the Unfinished Stair Timer . Or destroy the bell with a memory of the Astroegro's fall (lose a cherished memory permanently). Or give the bell to the Synod (gain a Synod Indulgence) or to the Ninth's Coterie (gain the Ninth's Annotation).

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power},breakable]
In Thepyrgos, power flows through elevation, education, and the precise application of law---but also through the hidden architecture of the Unfinished Stair. The Archon maintains authority through a complex web of guilds, synods, and civic institutions, but the real power lies with whoever can climb the stairs that should not exist. Social mobility is literally upward mobility, with stair-rights, bellmarks, and archive access serving as both currency and status symbols. But the highest status belongs to those who have seen the top of the Unfinished Stair---and remembered how to return.

Power flows through three overlapping hierarchies:
\begin{enumerate}
\item \textbf{Height:} The higher you climb, the more authority you hold. The Archon lives at the peak. The poor live in the lower tiers. Every stair is a status symbol.
\item \textbf{Knowledge:} The Library of Keys holds the city's memory. Those who control the archives control the past---and the future.
\item \textbf{The Unfinished:} The stairs that should not exist. Those who can climb them, who can ask the questions the Synod cannot answer, hold power beyond law.
\end{enumerate}

\textbf{Rewards are:}
\begin{itemize}
\item Keys (literal and metaphorical)
\item Stair-rights (permission to climb)
\item Bellmarks (authority made audible)
\item Archive access (knowledge as leverage)
\item A single climb of the Unfinished Stair (a question answered, a truth revealed, a vault opened)
\end{itemize}

\textbf{Costs are:}
\begin{itemize}
\item Memory (forgotten facts, lost names)
\item Obligation to a faction (the Witness, the Ninth, the Synod, or the Commissariat)
\item Permanent marks (scars, brands, missing steps)
\end{itemize}

The Lethai-thora, as the city's most respected archivists, hold a special kind of power: they know which shelves are watched, which keys are false, and which doors lead to the Ninth's vaults. Some whisper that the Archon is secretly a Lethai-thora puppet; others say the real ruler is the ghost of an Astroegro emperor, still counting stairs in the deep.

\textbf{For the GM:}  
Patronage in Thepyrgos revolves around access---to towers, to knowledge, to the right to climb, and to the hidden stairs. Rewards often take the form of keys, passes, or legal exemptions that can be leveraged into greater influence. To emphasize the metaphysical war:
\begin{itemize}
\item Tie rewards to visible symbols (tokens, keys, papers) that can be challenged, stolen, or voided.
\item Let rival guilds and the two Patrons (Witness and Ninth) issue conflicting rights, forcing players to choose whose favor matters more, and whose curse is worse.
\item Use the synods, towers, and staircases as arenas for social contests, where knowledge of law and custom determines success---but where the Ninth's silence and the Witness's memory can tip the scales.
\item Introduce the Unfinished Stair timer as a campaign tension; every time the party uses a hidden stair, they advance the timer and bring the Ninth's attention closer.
\end{itemize}
In Thepyrgos, your ascent is your standing, and your standing determines whether you rise or fall---or whether you are simply \emph{forgotten}.
\end{tcolorbox}

% Index entries
\index{Thepyrgos|see{City of Unfinished Proofs}}
\index{Theme generator}
\index{City of Unfinished Proofs generator}
\index{Places!Thepyrgos}
\index{People!Thepyrgos}
\index{Complications!Thepyrgos}
\index{Rewards!Thepyrgos}
\index{Archon}
\index{Matriarch of the Ladder}
\index{Lighthouse-Patriarch}
\index{Chain-Lantern order}
\index{Witch-Hunter Commissariat}
\index{Hyro people}
\index{Astroegro Empire}
\index{Khemesh}
\index{Lethai-thora}
\index{Blue Cistern}
\index{Beacon Crown}
\index{Library of Keys}
\index{Synod Hall}
\index{College of Ladders \& Lines}
\index{College of Bells \& Beacons}
\index{College of Waters \& Weights}
\index{College of Names \& Oaths}
\index{College of Mirrors \& Haze}
\index{College of Quiet}
\index{Saint Methodia}
\index{Cog-Cults}
\index{Daughters of the Cord}
\index{Savage Heart}
\index{Unfinished Stair}
\index{Ninth (Patron)}
\index{Witness (Patron)}
\index{Astroegro Vaults}
\index{Zakov, connection to Astroegro}
\index{Archon's Balance (timer)}
\index{Commissariat Pressure (timer)}
\index{Ninth's Progress (timer)}
\index{Witness's Vigil (timer)}
\index{Faction Relationship Dial!Witness vs Ninth}