\section{Tulkani --- ``The Itinerant Courts''}
\label{chap:tulkani}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The knot remembers. The story answers. The circle holds.}

\begin{quote}
  \textbf{Caravan Judge (Elite):}\\
  ``The road does not forgive a broken promise. It only reschedules the telling. A vow spoken under the open sky is witnessed by the wind, the grass, and the wheels of every cart that passes. I do not write judgments. I tie knots. The knots remember. When you are ready to share, untie one. If you are not ready, tie another. The circle waits.''
\end{quote}

\begin{quote}
  \textbf{Wagon-Keeper (Commoner):}\\
  ``My wagon is my hearth, my court, my grave. The Hollow cannot cross the circle of wheels unless invited. The collector cannot enter without a story. The lord cannot claim his due unless we camp on his land for nine nights. On the eighth night, we leave. That is the law --- not because we fear the lord, but because the ninth night belongs to someone else.''
\end{quote}

\noindent\textbf{Genre / Mood:} Itinerant community, story-driven kinship, folk horror at the edge of the Hollow.

\noindent\textbf{Starting Location:} A wagon circle at a crossroads, where nine painted wagons form a ring. A Caravan Judge sits on an overturned barrel, a knotted cord in her hand. ``You are standing on ground that belongs to the lord of the eastern valley. He will be here at dawn to press his claim. But you are in the circle. You are safe until the ninth hour. Tell me a story --- a true one --- and I will tell you how to leave before the lord arrives. And if you hear a whisper from the Dream-Wagon, do not answer. That is not a voice for waking ears.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Tulkani have no fixed homeland; they are wanderers who follow the seasons and the trade routes. Their wagons are their hearths, their circles are their courts, and their knots are their memories. They are welcome everywhere and trusted nowhere --- a position they wear with wry humor. A Tulkani caravan can appear at a lord's border, camp for seven nights, and leave before the eighth --- because on the eighth night, the lord can demand rent. On the ninth night, something else comes.

Their courts are circles of wagons; their judges are elders who carry braided cords. Each knot is a vow, a story, or a memory shared --- but never merely a record. They are the keepers of an older rhythm, one that predates the crowns and the tariffs. Their neutrality is sacred: a Tulkani camp is sanctuary for any traveler, regardless of allegiance, as long as they respect the circle and offer a story when asked.

\vspace{2mm}
\textbf{What you notice first:} The smell of woodsmoke and cooking oil. The painted wheels of wagons arranged in a circle, each wheel a different color. The way everyone touches a knotted cord before speaking of a journey.

\textbf{The one rule that kills newcomers:} Never sleep outside the circle at night. The wagon ring is sanctuary. Outside it, the Hollow walks. A Tulkani will not rescue you --- they will only mourn you at dawn.
\end{tcolorbox}

\subsection*{The Curse of the Ninth Night: Why We Wander}
\label{subsec:tulkani-origin}

\begin{tcolorbox}[colback=black!2,colframe=red!70!black,title={A Fragment of the Dream-Cord of the First Night (as sung by the Caravan Judges)}]
\textit{``The Spider spun the net, the Snake shed the truth, the Raven closed the sky. But the one who dreamed --- she whispered through the cracks. `Take these threads,' she said. `Tie them in eights. The ninth is for me. I will sleep, and you will carry my waking. Never stay nine nights in the same breath. The Hollow counts. The Hollow follows. If it catches you, give it a story --- not silver, not blood. A story it has never heard. That is the toll. That is the mercy.'''}
\end{tcolorbox}

No Tulkani speaks openly of the Dreamer. The old ones say that before the first caravan, before the roads were named, three sisters watched the world burn and rebuilt it from ash. Two of them decided that mortals must never know the fire. The third --- the one who dreamed of a different path --- was sealed into a sleep that would last as long as the mountains. But before she closed her eyes, she taught a handful of herders and smiths how to knot her dreams into cord. ``You will be my memory,'' she told them. ``And the Hollow will be my shadow. Keep moving, or it will ask you what I whispered.''

Thus the Tulkani became wanderers. They do not stay nine nights because on the ninth night, the Hollow catches up. It does not attack. It asks for a story. If the Tulkani can tell one that the Dreamer herself never heard, the Hollow retreats --- until the next camp, the next ninth night. If they cannot, it takes a knot from the judge's cord, and a piece of the Dreamer's memory fades forever.

The Tulkani worship Kuva, the sky father, openly. They light no shrines to the Dreamer. But every circle has a \textbf{Dream-Wagon} --- a painted wagon with no wheels, its door a curtain of knots. Inside, a single candle burns beside a bed of straw. The old ones say that if you listen at midnight, you can hear the Dreamer breathing. They also say that the Raven's shadow falls across the wagon on the ninth night of every ninth season. No one has ever seen it and remained in the circle.

\subsection*{Regional Mood: The Measure of the Knot}

Power here is measured in memory-cords, shared stories, and the number of nights camped on a lord's land. The Tulkani have no king, no judges, no written law --- only the Caravan Circle, where disputes are settled by storytelling. A liar is exposed not by evidence but by the knots on his cord: if a knot is tied in false witness, it will tighten until blood is drawn. The Dreamer's cult is not a religion; it is a survival tactic --- the dream inside the wagon, the whisper in the creaking wheel. The Tulkani give no permanent coin, but they answer in stories. A lord who demands silver will be told a story so boring that his crops fail. A lord who offers hospitality will be told a story so beautiful that his harvest doubles. The Tulkani do not keep accounts; they keep faith in reciprocity. What you give, you get back --- maybe tomorrow, maybe in the next valley. But the circle remembers.

\subsection*{Prominent NPCs of the Tulkani}
\label{sec:tulkani-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Elder Sarnai & Caravan Judge of the Nine Knots & Keep the circle whole and the Dreamer's memory intact & She ties a knot for every judgment; her cord drags on the ground. When it touches dirt, she whispers a name. \\
\hline
Temur & Wagon-Keeper & Find new grazing ground before the ninth night & He paints a new color on his wagon each season; the colors are memories of places he has never seen. \\
\hline
Gulshat & Story-Seller & Spread the dream through tales & She carries a dream-cord with nine knots; the ninth knot is loose. She never ties it. \\
\hline
The Knife (Unnamed) & Ikasha's silent agent & Free a spirit bound by the Veiled Sisters & He wears a mask made of knotted cord; underneath, he has no face --- only a memory of a memory. \\
\hline
Vow-Broken & An exile who cut his own cord & Earn back his place in the circle & He carries a single knot tied in his own hair; the knot is his promise to return. His shadow is missing two fingers. \\
\hline
\end{longtable}

\subsection*{Names of the Tulkani}
\label{sec:tulkani-names}

Inspired by the nomadic traditions of the steppe and desert: short, guttural, with honorifics and clan prefixes. Surnames are often replaced by epithets describing a deed or a knot.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Clan/Epithet} & \textbf{Knot Name} \\
\hline
Temur & Sarnai & the Wagon-Keeper & Ninth Cord \\
\hline
Batu & Gulshat & the Story-Seller & Dream-Knot \\
\hline
Orhan & Nergui & the Knife & Unbound \\
\hline
Cengiz & Altan & Vow-Broken & Cut Cord \\
\hline
Toghrul & Odval & Circle-Keeper & Neutral Knot \\
\hline
\end{longtable}

\paragraph*{(Circle/Wagon/Ground)} A wagon circle at dusk, lanterns lit; a painted wagon with a dream-cord hanging from the wheel; a crossroads where nine roads meet.

\section*{Spades --- Places}
\label{sec:tulkani-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Wagon Circle (Night)} --- Nine wagons arranged in a ring, lanterns burning, wheels touching.
  \hfill \textit{Inside the circle, the Hollow cannot enter. Outside, it waits. The lanterns must never go out. On the ninth night, one flickers.} [SANCTUARY]
\item \textbf{Caravan Court (Day)} --- A carpet spread on grass; a judge on a barrel; a cord on the ground.
  \hfill \textit{The cord is the circle's memory. Step over it, and you accept the judge's telling.} [JUSTICE]
\item \textbf{Dream-Wagon (Ikasha's Shrine)} --- A painted wagon with no wheels, sitting on stones. The door is a curtain of knots.
  \hfill \textit{Inside, a single candle and a bed of straw. The Dreamer sleeps there. Do not wake her.} [MYSTIC]
\item \textbf{Cord-Maker's Stall} --- A canvas awning where a woman braids horsehair and silk.
  \hfill \textit{Each cord is a memory. The color tells the weight: white for truth, red for a promise, black for a secret never spoken.} [CRAFT]
\item \textbf{Crossroads of the Ninth Road} --- A meeting of nine paths, only eight of which are visible.
  \hfill \textit{The ninth road is for the Dreamer's shadow. The Tulkani leave an offering of a single knotted thread there each full moon.} [THRESHOLD]
\item \textbf{Abandoned Circle} --- A ring of burnt grass and wheel ruts. No wagons. No people.
  \hfill \textit{The circle was broken on the ninth night. The Hollow came. The grass will not grow back.} [OMEN]
\item \textbf{Memory-Cord Tree} --- An old oak with hundreds of knotted cords hanging from its branches.
  \hfill \textit{Each cord is a promise kept --- or a story shared. The tree remembers. The wind unties one each season.} [COMMUNITY]
\item \textbf{Hollow's Gate (Seasonal)} --- A place where the boundary is thin, marked by a single painted stone.
  \hfill \textit{The Tulkani camp here only on the new moon. On the full moon, they are a day's ride away. The stone is warm.} [THRESHOLD]
\item \textbf{Tinkanar (Itinerant Caravanserai)} --- A mobile market of wagons, open for one day each season.
  \hfill \textit{The location changes. The Tulkani announce it by dream --- the dream of a sleeping merchant.} [TRADE]
\item[J] \textbf{Story-Well} --- A dry well where the Tulkani throw written stories on parchment.
  \hfill \textit{The stories are offerings to the Dreamer. The well is full. The parchment never rots. At midnight, you can hear pages turning.} [RITUAL]
\item[Q] \textbf{The Ninth Cord's Post} --- A wooden post at the centre of every circle, with nine notches.
  \hfill \textit{The ninth notch is never carved. It is left for the Hollow. The Hollow never takes it --- only counts it.} [SYMBOL]
\item[K] \textbf{The Wandering Hearth} --- A wagon that is never hitched, always moving. It carries the council fire.
  \hfill \textit{The fire has been burning for three hundred years. It has never touched the ground. The ashes are knotted.} [LEGEND]
\item[A] \textbf{Ikasha's Dreaming Place (Myth)} --- A cave beneath a hill, where She Who Sleeps lies on a bed of knotted cords.
  \hfill \textit{If you untie a cord, she wakes. If she wakes, the world ends --- or begins again. No one knows which.} [MYTH]
\end{enumerate}

\paragraph*{(Judge/Keeper/Storyteller)} Caravan Judge with a dragging cord; Wagon-Keeper painting a wheel; Story-Seller with a dream-cord and a cup of water.

\section*{Hearts --- People \& Factions}
\label{sec:tulkani-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Caravan Judge} --- Old, gray-braided, a cord so long it drags on the ground.
  \hfill \textit{She ties a knot for every judgment. Untying a knot voids the telling. She has never untied one. The cord remembers what she does not.} [JUSTICE]
\item \textbf{Wagon-Keeper} --- A woman with paint on her hands, a brush behind her ear.
  \hfill \textit{She paints a new color on her wagon each season. The colors are memories of places her grandmother saw. She is running out of colors.} [CRAFT]
\item \textbf{Story-Seller (Gulshat)} --- A young woman with a dream-cord around her wrist. She speaks in rhymes.
  \hfill \textit{Her stories are true --- but only if you listen with your eyes closed. If you open them, the ending changes.} [SOCIAL] [MYSTIC]
\item \textbf{Ikasha's Knife} --- A hooded figure with a mask of knotted cord. No face underneath.
  \hfill \textit{He cuts the cords of those who break the circle's trust. The cut cord becomes a ghost. The ghost follows the breaker, whispering the story they tried to forget.} [JUSTICE] [DEATH]
\item \textbf{Vow-Broken (Exile)} --- A man with a single knot tied in his own hair. He never smiles.
  \hfill \textit{He is saving stories to buy back his name. He needs nine. He has eight. The ninth is the one he cannot tell.} [COMMUNITY]
\item \textbf{Cord-Maker} --- A woman with braids of horsehair and silk between her fingers.
  \hfill \textit{She can tell the meaning of a promise by the feel of the hair. Horsehair is heavy. Silk is light. A knotted hair from a sleeping child is heaviest of all.} [CRAFT]
\item \textbf{Hollow-Watcher} --- A boy who sits at the edge of the circle at dusk, watching the dark.
  \hfill \textit{He can see the Hollow when it walks. He points. The adults throw salt. The salt is from a pan that dried up nine generations ago.} [SCOUT]
\item \textbf{The Dreamer} --- A woman who sleeps in the Dream-Wagon. She never wakes during the day.
  \hfill \textit{She dreams the caravan's path. The wagon follows her dreams. When she smiles, the road is easy. When she weeps, the Hollow is close.} [MYSTIC]
\item \textbf{Tinkanar Merchant} --- A fat man with nine bells on his hat. He sells knots and stories.
  \hfill \textit{He accepts only promises in trade. His reckoning is a cord. It has nine thousand knots. The oldest is tied around a memory of the Dreamer's first whisper.} [TRADE]
\item[J] \textbf{The Ghost of the Ninth Night} --- A translucent figure seen at abandoned circles.
  \hfill \textit{It was a Tulkani who stayed too long. It asks: ``Is it the ninth hour yet?'' Do not answer. It is not asking you.} [DEATH]
\item[Q] \textbf{The First Wagon-Keeper (Legend)} --- A skeleton driving a wagon of bones. The wheels are skulls.
  \hfill \textit{It appears when a circle is broken. It offers a ride. The ride leads to the Hollow --- or to the Dreamer's cave. No one has returned to say which.} [LEGEND]
\item[K] \textbf{Ikasha's Voice} --- A whisper heard in the Dream-Wagon at midnight.
  \hfill \textit{It speaks in riddles. ``The ninth cord is not a cord,'' it says. ``It is the space between the knots. Count carefully. The Hollow does.''} [MYSTIC]
\item[A] \textbf{The Unwitnessed} --- A Tulkani who never tied a single knot. No one remembers her name.
  \hfill \textit{She walks through circles unnoticed. She can break any vow because no one saw her make it. She is the Dreamer's memory of herself, before the sealing.} [MYTH]
\end{enumerate}

\paragraph*{(Toll/Story/Hollow)} A lord expects his due for camping on his land --- negotiate nights or tell a story; the Hollow walks on the ninth night --- a character hears their own name whispered; a stranger recognizes your memory-cord and asks for a tale.

\section*{Clubs --- Complications/Threats}
\label{sec:tulkani-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Recognition (Memory-Cord Seen)} --- Someone from your past sees the knots you carry. They ask for a story you promised.
  \hfill \textit{The story must be about a promise you kept or broke. If you cannot tell one, they tie a new knot --- and the Hollow watches.} [SOCIAL] [RECIPROCITY]
\item \textbf{Lord's Claim} --- The local lord arrives at the circle at dawn. ``You have been here seven nights. Now comes the eighth.''
  \hfill \textit{The price is a story that flatters him. If the story fails, he seizes a wagon. The circle holds its breath --- and the Hollow listens.} [AUTHORITY] [NEGOTIATION]
\item \textbf{Hollow Walks (Ninth Night)} --- Someone in the circle has miscounted. The ninth hour approaches.
  \hfill \textit{The lanterns flicker. A shadow stands at the edge. It whispers a name. The name is the last person who broke an oath --- or the next.} [DEATH] [FEAR]
\item \textbf{Memory-Cord Tangled} --- Two promises conflict. The knots tighten. Blood is drawn.
  \hfill \textit{The judge must cut one cord. The severed cord becomes a ghost. The ghost accuses the cutter of favoritism --- and the Hollow listens.} [CONFLICT] [RECIPROCITY]
\item \textbf{Wagon Wheel Broken} --- A wheel cracks. The circle is incomplete. The Hollow can enter.
  \hfill \textit{The wheel must be repaired before dusk. The repair requires a story --- a true one told to the wood. The wood listens. The wood remembers.} [DISASTER]
\item \textbf{Tulkani Schism} --- Two judges dispute the meaning of a knot. The circle splits.
  \hfill \textit{Each judge claims authority. The party must choose which cord to follow. The other cord will remember --- and so will the Hollow.} [POLITICS]
\item \textbf{Ikasha's Dream Disturbed} --- A stranger enters the Dream-Wagon. The Dreamer wakes.
  \hfill \textit{Ikasha stirs. The ground trembles. The stranger must tell a story so good that the Dreamer falls back asleep. The stranger is you.} [MYSTIC] [CRISIS]
\item \textbf{Story Crossing (Bridge/Pass)} --- A guard demands a story to cross. Your story is not good enough.
  \hfill \textit{The guard demands a truth instead. The truth must be a secret. The secret must hurt. The guard will remember --- and the Hollow will too.} [TRAVEL] [SOCIAL]
\item \textbf{Vow-Broken's Return} --- An exile tries to rejoin the circle. The judge refuses.
  \hfill \textit{The exile demands the party speak for him. If they refuse, he cuts his own throat. If they speak, they hold a knot --- and a story.} [COMMUNITY] [DRAMA]
\item[J] \textbf{The Ninth Cord Cut} --- Someone cuts the ninth knot on the central post. The Hollow floods in.
  \hfill \textit{The lanterns go out. The circle becomes a cage. The party must re-tie the knot in the dark. The cord is warm. It has a pulse.} [DEATH] [TRAP]
\item[Q] \textbf{Ikasha's Knife Betrayed} --- The Knife is offered a bribe to cut a false cord. He refuses. The briber attacks.
  \hfill \textit{The Knife is wounded. He asks the party to carry his mask. The mask weighs nothing until you put it on. Then it weighs everything --- every knot ever cut.} [COMBAT] [JUSTICE]
\item[K] \textbf{Tinkanar Closed (No Stories)} --- The mobile market is empty. The merchant has run out of knots.
  \hfill \textit{He is selling his own name. The price is a story that makes him weep. He has not wept in forty years. The last story he heard was a lie. The Hollow is waiting.} [TRADE] [CRISIS]
\item[A] \textbf{Ikasha Wakes} --- The Dreamer dies. Ikasha opens her eyes. The sky turns white.
  \hfill \textit{She asks one question: ``Who cut the ninth cord?'' The answer must be a name. If no name is given, she takes all names --- and the Hollow follows her out.} [APOCALYPSE]
\end{enumerate}

\paragraph*{(Cord/Bead/Charm)} Memory-bead (a single knot, can be used to call in a story once); safe-haven oath (a knotted cord that guarantees sanctuary in any Tulkani circle); dream-cord (gift from Ikasha --- one question answered in a dream, but the ninth knot tightens).

\section*{Diamonds --- Rewards/Leverage}
\label{sec:tulkani-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Memory-Bead (Single Knot)} --- A small knot on a leather thong. Present it to ask for a story from any Tulkani.
  \hfill \textit{The knot will loosen after use. If it is not retied within a year, the story becomes a burden --- and the Hollow will ask for it.} [SOCIAL]
\item \textbf{Safe-Haven Oath (Knotted Cord)} --- A braided cord with three knots. Hang it on your wagon to guarantee sanctuary in any Tulkani circle.
  \hfill \textit{The cord must be renewed each season. A new knot is added. After nine knots, it becomes a dream-cord --- and the Hollow will notice you.} [TRAVEL]
\item \textbf{Dream-Cord (Ikasha's Gift)} --- A silk cord with nine knots, the ninth loose. Once, sleep on it to ask Ikasha one question.
  \hfill \textit{She will answer in a dream. The answer is always a riddle. The ninth knot tightens after use. When it tightens fully, Ikasha wakes --- and the world changes.} [MYSTIC]
\item \textbf{Caravan Judge's Cord (Borrowed)} --- A length of braided horsehair. Tie it around your wrist to gain $+1$~Position in any negotiation with Tulkani.
  \hfill \textit{The cord is the judge's own. She will want it back. The return is a negotiation --- and a story.} [SOCIAL]
\item \textbf{Wagon-Keeper's Paint (Pot)} --- A clay pot of blue pigment. Paint your own wagon wheel, and no lord will demand rent for nine days.
  \hfill \textit{The paint is made from crushed sky-stone. It glows under moonlight. The Hollow avoids it --- but the lord will send a spy.} [TRAVEL]
\item \textbf{Story-Seller's Verse} --- A single verse of a true story. Recite it to calm a spirit or end a feud.
  \hfill \textit{The verse works once. After that, the story belongs to the listener. You cannot tell it again --- but you can tell a different truth.} [SOCIAL] [MYSTIC]
\item \textbf{Ikasha's Mask (Knotted)} --- A mask of knotted cord. Wear it to become invisible to the Hollow for one night.
  \hfill \textit{The mask has no eye holes. You must walk by memory. The memory may fail. If it fails, the Hollow sees you --- and smiles.} [STEALTH] [DEATH]
\item \textbf{Hollow-Watcher's Salt Pouch} --- A leather pouch of salt blessed by the boy who watches the circle. Throw it to banish an Oath-Keeper.
  \hfill \textit{The salt is bitter. The Oath-Keeper will taste it. It will remember your face --- and your smell.} [COMBAT] [DEATH]
\item \textbf{Tinkanar Merchant's Bell} --- One of nine bells from the merchant's hat. Ring it to summon him anywhere.
  \hfill \textit{He will come. He will demand payment in stories. He will not leave until paid. He has a long memory --- and the Hollow listens.} [TRADE]
\item[J] \textbf{Vow-Broken's Hair Knot} --- A single knot tied in human hair. It contains a broken promise. Use it to void a contract.
  \hfill \textit{The knot is from an exile. He wants it back. He will give a story in return --- the story of how he broke his vow.} [SOCIAL] [MAGIC]
\item[Q] \textbf{The Ninth Cord's Post Fragment} --- A splinter of wood from the central post. Touch it to any knot to learn who tied it.
  \hfill \textit{The splinter is warm. It will burn your hand if the knot was tied in false witness. The truth leaves no scar --- only a memory.} [MYSTIC]
\item[K] \textbf{First Wagon-Keeper's Skull} --- A human skull painted with wheel patterns. Place it at the centre of a circle to keep the Hollow away for nine nights.
  \hfill \textit{The skull whispers on the ninth night. It asks: ``Is it time?'' Say no. If you say yes, it will take your place --- and your story.} [DEATH] [SANCTUARY]
\item[A] \textbf{Ikasha's Dream (Bottle)} --- A glass bottle containing a sleeping breath. Break it to send a question to Ikasha.
  \hfill \textit{She will answer in a vision. The vision lasts one minute. The cost is one memory. The memory is forgotten forever. The memory is always a happy one --- until it is gone.} [MYSTIC]
\end{enumerate}

\section*{Quick use notes}
\label{sec:tulkani-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer ($2$--$5$ $\rightarrow$ $4$, $6$--$10$ $\rightarrow$ $6$, J/Q/K $\rightarrow$ $8$, A $\rightarrow$ $10$).
\item Diamonds are codified outcomes (knots, cords, beads) that change position rather than call for a roll.
\item If any A appears, echo \textbf{knot-omens} --- a cord that tightens without being pulled, a knot that unties itself, a wheel that stops turning at the ninth spoke. The Hollow is close.
\end{itemize}

\subsection*{Additional Features: Itinerant Mechanics}
\label{sec:tulkani-mechanics}

\begin{itemize}
\item \textbf{The Ninth Night Rule:} A Tulkani caravan may camp on a lord's land for seven nights without payment. On the eighth night, the lord may ask his due --- typically a story, a promise, or a small service. On the ninth night, the Hollow comes. No Tulkani stays nine nights in one place. A party camping with Tulkani must leave by the eighth dusk or face the Hollow's attention.
  \hfill \textit{Mechanic:} Each night the party camps in a Tulkani circle, mark a segment on a \textbf{Ninth Night Timer [$8$]}. At $7$ segments, the lord (or local authority) arrives to ask his due. At $8$ segments, the Hollow manifests. The party can leave before $8$ to reset the timer.
\item \textbf{Story as Kinship:} Tulkani do not use coin. What passes between them are stories, promises, and memories. A good story can buy a meal, a horse, or safe passage. A bad story can start a feud. The GM should encourage players to tell brief in-character stories; the quality (or a Presence + Sway roll) determines its worth. On a success, the story is worth $1$~Measure of value; on a critical success, $2$. On a failure, the listener is offended --- the party gains $-1$~Position on their next social roll in that circle. A player may also offer a truth (a secret, a confession, a memory) instead of a story; a truth is always worth $2$~Measures but marks $1$~Bond to the Dreamer.
\item \textbf{Land Rent and Neutrality:} Tulkani circles are neutral ground. No lord may collect taxes or enforce laws inside a circle, regardless of whose territory it sits on. However, the lord may ask his due at the circle's edge, and the Tulkani will negotiate. A party inside a Tulkani circle is under the Caravan Judge's protection --- but if the party commits a crime, the judge may expel them to the Hollow.
\item \textbf{Memory-Cords and Reciprocity:} Every Tulkani carries a memory-cord --- a braided strand of horsehair and silk, with knots tied for every promise made and kept. A Tulkani who breaks a promise must cut the corresponding knot; the cut cord becomes a ghost that follows the oath-breaker until the promise is fulfilled. Outsiders may borrow a Tulkani memory-cord; the knots then record their promises. The Tulkani see this as reciprocity: what is given returns, and the circle always balances.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Unwitnessed Weaver (Inaea):} The knot that binds gently; the cord that remembers mercy.
    \item \textbf{The Cord-Cutter (Isoka):} The clean break; the shedding of false promises.
    \item \textbf{The Dreamer Below (Ikasha):} She who sleeps in the wagon; she who dreams the road.
    \item \textbf{The Ninth Traveler (The Traveler):} The road that has no end; the threshold that is always one step away.
  \end{itemize}
\end{itemize}

\subsection*{The Ninth Night: A Story Timer }
\label{subsec:tulkani-ninth-night}

When the Hollow walks on the ninth night (or when the \textbf{Ninth Night Timer } reaches $8$), it does not attack. It \emph{asks}. The GM should select or roll on the table below to determine what story the Hollow demands. The party must then tell a story (in character, as a group) that satisfies the Hollow's request. The GM judges success based on creativity, coherence, and whether the story fits the prompt.

\begin{longtable}{|p{0.2\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{d8} & \textbf{The Hollow Asks For...} \\
\hline
1 & A story about a door that was left open. \\
\hline
2 & A tale of a promise kept when breaking it would have been easier. \\
\hline
3 & A confession of a small cruelty you have never admitted. \\
\hline
4 & A story that ends with the number eight, not nine. \\
\hline
5 & A memory of a road you will never walk again. \\
\hline
6 & A lie that became true because someone believed it. \\
\hline
7 & A story about a gift you gave and never received thanks for. \\
\hline
8 & A tale of a name you have forgotten --- and why you forgot it. \\
\hline
\end{longtable}

If the party succeeds (the GM decides), the Hollow retreats, and the \textbf{Ninth Night Timer } resets to $0$. If they fail, the Hollow takes a knot --- one of the party's active Boons, a point of Reputation, or a memory (erase a minor fact or a Bond for one session). The timer resets but the Hollow's attention lingers (the GM gains $2$~SB to spend on future supernatural complications).

\subsection*{Story Bonds: Promises and Stakes}
\label{subsec:tulkani-story-economy}

The Tulkani treat stories as the most valued exchange. When dealing with Tulkani, the party may offer a \textbf{Story Promise} instead of coin or goods. A Story Promise is a verbal contract: the party agrees to tell a true story (on a future occasion) to the Tulkani elder or to a specific person the elder names. The promise is witnessed by the circle --- a knot is tied on a cord.

\begin{itemize}
\item \textbf{Stakes:} The Tulkani may ask for something held in trust --- a minor secret, a lock of hair, or a written name --- set against the promise. If the promise is broken, the stake is given to the Hollow (it vanishes, and the party gains $+2$ to the \textbf{Hollow Attention} timer).
\item \textbf{Echo:} A story promised but delayed echoes. Each season the promise goes unfulfilled, the GM may ask the party to tell a \emph{small} story (a paragraph) to the table as a reminder. If they refuse, the Hollow's attention ticks $+1$.
\item \textbf{Forgiveness:} A story promise can be forgiven if the party tells a story that makes the Caravan Judge weep --- or laugh so hard she cuts the cord herself. This is a high-DC social roll (typically Presence + Performance vs.\ DV~$5$) and may require an actual in-character performance.
\end{itemize}

\subsection*{Mini-Generator: Stories the Tulkani Trade}
\label{subsec:tulkani-story-generator}

When a Tulkani offers a story in exchange for information, goods, or passage, roll on this table to determine its nature and mechanical effect.

\begin{longtable}{|p{0.2\textwidth}|p{0.6\textwidth}|}
\hline
\textbf{d6} & \textbf{Story Type \& Effect} \\
\hline
1 & \textbf{A Warning.} The story reveals a hidden danger (the GM must tell the party about one upcoming hazard or ambush). The story is true --- but the teller may have exaggerated. Effect: The party gains $+1$~Position to avoid that specific hazard. \\
\hline
2 & \textbf{A Route.} The story describes a forgotten path, a smuggler's trail, or a safe ford. Effect: For the next travel leg, the party may ignore one environmental penalty (difficult terrain, bad weather, etc.). \\
\hline
3 & \textbf{A Claim.} The story recounts an old promise tied to the teller's family. The party may choose to honor it (gain trust from the Tulkani but also a \textbf{Mark} --- a minor promise that the GM can call in later) or ignore it (the Tulkani is offended; $-1$~Position on future social rolls with that circle). \\
\hline
4 & \textbf{A Secret.} The story reveals a piece of gossip about a local lord, merchant, or rival faction. Effect: The party gains a one-use \textbf{Secret Leverage} --- $+2$~dice on a single social roll against that person. \\
\hline
5 & \textbf{A Lament.} The story is a tragedy --- a death, a lost love, a failed quest. Those who listen properly gain $+1$~die to resist fear or despair for the next scene. The teller, however, becomes melancholy ($-1$~die on their next roll). \\
\hline
6 & \textbf{A Dream.} The story is not quite true --- it is a dream the Dreamer had. The party may ask the GM one yes/no question about the near future (the next scene or the next leg of travel). The answer is always cryptic, but never false. \\
\hline
\end{longtable}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\label{sec:tulkani-lore}

\begin{itemize}
  \item \textbf{Tulkani Memory-Cords and the Sidhi Freedman's Seal:} Sidhi merchants on the Brass Coast accept Tulkani memory-cords as proof of standing. A well-knotted cord can buy a ship. A cut cord can sink one. The Sidhi know that a Tulkani knot is worth more than a signature --- and that the Hollow reads over their shoulder.
  \item \textbf{Ikasha and the Veiled Sisters (Ashaan):} The Tulkani are the primary spreaders of Ikasha's cult outside Ashaan. A Breath-less agent may be disguised as a Tulkani story-seller; a Tulkani caravan may be a mobile safe house. The Sisters have tried to infiltrate the circles three times. Each time, the spy was found with a cord around her throat --- tied in a knot she could not untie.
  \item \textbf{The Ninth Night and the Hollow (Eastern Realms):} The Tulkani's ninth-night rule is the only reason the Hollow has not overrun the eastern trade routes. Aelaerem hedge-keepers sometimes hire Tulkani guides to reset the Hollow's attention timers. The payment is always a story --- and the story is always about a door that was left open.
  \item \textbf{Story Debts and the Ukwe (Sekogo):} The Tulkani believe that the Ukwe is the first memory-cord tree. They leave offerings of knotted grass at its roots. A Tulkani judge may consult the Ukwe to verify a story's truth. The Ukwe speaks in rustles. A single rustle means true. Two mean false. Three mean the story is not finished --- and the Hollow is listening.
\end{itemize}

\subsection*{Tulkani Superstitions (burn for $+1$~die once)}
\begin{itemize}
  \item \textbf{Bread to the Circle:} Crumble a crust at the centre of the wagon ring. Ignore the first \emph{Recognition} penalty. [SUPERSTITION]
  \item \textbf{Knot the Lantern:} Tie a single knot in the rope of a lantern. Cancel \emph{Lord's Claim} or take $-1$~die on your next negotiation (your choice). [SUPERSTITION]
  \item \textbf{Name to the Ninth Post:} Whisper a dead judge's name into the central post. Gain $+1$~Effect when telling a story this scene, but mark $1$~Bond to the Dreamer. [SUPERSTITION]
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:tulkani-sb}

\paragraph{Minor Complications ($1$~SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from Tulkani judges or local lords.
\item \textbf{Noise:} Sounds of your actions alert nearby circle-watchers or Hollow scouts.
\item \textbf{Trace:} Evidence of your passage marks your route for vow-breakers or story-hunters.
\item \textbf{Delay:} A brief but meaningful setback costs you time or favorable camping ground.
\item \textbf{Supply Strain:} Mark $+1$~segment on a relevant resource timer.
\end{itemize}

\paragraph{Moderate Setbacks ($2$~SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} Local Caravan Judge or lord's herald becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to lantern failure or broken wheel.
\item \textbf{Foe Appears:} A story-hunter or Hollow manifestation arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes Compromised/Neglected.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble ($3$~SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional Tulkani riders, lord's men, or Hollow spirits arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes --- knot cut / wagon broken / ninth night triggered.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by $1$~segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue $1$/Harm $1$/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns ($4+$~SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{High Caravan Judge, Ikasha's Knife, or Hollow's Oath-Keeper} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically --- \textbf{circle breaks / wagon burns / lanterns go out}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice --- \textbf{omen appears / blessing lost / curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Tulkani (Knot Law):} Cords tighten without warning, a knot unties itself, a wheel stops at the ninth spoke.
\item \textbf{Tulkani (Story Echo):} A story you told returns to you changed; a listener laughs at a tragedy; the wind carries your words to someone who should not hear.
\item \textbf{Tulkani (Ninth Night):} A lantern flickers, a shadow stands at the circle's edge, a whisper says your name from the dark.
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Stealth):} On a hit, shift one band. On a strong hit, also impose \emph{False Knot}.
\item \textbf{Cut the Cord (Presence + Sway):} On a hit, freeze range for one exchange; if a memory-bead is in your possession, $+1$~die.
\item \textbf{Weather the Hollow (Resolve + Spirit):} On a hit in \emph{Hollow Walks}, you pick who hears the whisper; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (Knot \& Dream)}
\begin{itemize}
\item \textbf{The Cord-Cutter (Nine Broken Promises)} --- \emph{Phases}: Memory-Cords Severed $\rightarrow$ Orchestrated Exile $\rightarrow$ Circle Coup. \emph{Cracks}: re-tie a cut cord, expose a false judge, survive a night outside the circle.
\item \textbf{The Dream-Eater (Ikasha's Shadow)} --- \emph{Phases}: Dreamer Waking $\rightarrow$ Procession of Empty Wagons $\rightarrow$ Knot-Baptism of a Storyteller. \emph{Cracks}: relic dream-cord, survivor testimony, the ninth cord refusing to tighten.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
Among the Tulkani, power flows through knots, stories, and the memory of the circle. A Caravan Judge's authority is recognized by every lord from the steppe to the coast --- not because of armies, but because the Hollow respects the judge's cord. The Tulkani have no temples, no kings, no borders. They have the ninth night, and that is enough.

\textbf{For the GM:} Power in Tulkani society is measured in knots tied, stories told, and nights counted. Patronage often takes the form of memory-beads, dream-cords, or safe-haven oaths --- each holding the spirit of a promise. To emphasize the itinerant mechanics:
\begin{itemize}
\item Require players to negotiate land rent with local lords, using stories or promises as their answer. A failed negotiation may lead to the Hollow.
\item Use the \textbf{Ninth Night Timer } as a tension mechanic: every night in a Tulkani camp advances the timer. At $8$, the Hollow manifests unless the party leaves.
\item Let memory-cords be physical props --- actual knotted strings that players can hold, tie, or cut. A cut cord is a dramatic moment.
\item Emphasize that storytelling is a skill: reward creative players with mechanical bonuses, and let failed stories have consequences --- a bored lord may tighten his claim, an offended spirit may curse.
\item Treat Tulkani society as a web of reciprocity, not a reckoning of accounts. What you give, you get back --- but the circle decides when.
\end{itemize}
In Tulkani society, the knot remembers everything --- and the ninth night forgets nothing.
\end{tcolorbox}

% Index entries
\index{Tulkani|see{Itinerant Courts}}
\index{Theme generator}
\index{Itinerant Courts generator}
\index{Places!Tulkani}
\index{People!Tulkani}
\index{Complications!Tulkani}
\index{Rewards!Tulkani}
\index{Caravan Judge}
\index{Dream-Wagon}
\index{Ikasha}
\index{Ninth Night}
\index{Memory-Cord}
\index{Story as Kinship}
\index{Hollow (attention timer)}
\index{Local Patrons!Tulkani}