\section{Valewood --- ``Empire Under Leaves''}
\label{chap:valewood}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The forest remembers. Speak truth, step lightly, and never eat the fruit of a tree with a face.}

\begin{quote}
  \textbf{Hazel Queen (Elite):}  
  ``The forest here remembers empires that crumbled to dust, and its laws are older than your kingdoms. To walk these paths is to enter into covenant with powers that measure time in seasons, not years.''
\end{quote}

\begin{quote}
  \textbf{Pathweaver (Commoner):}  
  ``The woods here got a mind of their own -- they'll lead you where you need to go if you ask right, but they'll just as soon lose you if you're carrying ill intent. Best know which is which before you wander off the path.''
\end{quote}

\textbf{Genre / Mood:} Dark forest, fey remnant, imperial echo, folk horror.

\textbf{Starting Location:} A moss‑covered amphitheatre where roots crawl through marble benches and owls whisper verdicts to the wind. An Owl‑Sister speaks in riddles: ``The Lethai‑ar stir. The paths remember your true name. Speak it into the hollow of this stone, or the forest will learn it from your footsteps.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Valewood is older than kingdoms, older than the maps that skirt its borders. Beneath its canopy lie the bones of a Lethai empire whose name is forgotten, its towers swallowed by vine and root, its star‑roads cracked and phasing between worlds. The air hums with half‑remembered songs of magic, and the ruins bleed their mysteries into root and stream. Fae courts linger in clearings that shift with moonlight, beast‑kin patrol moss‑choked cairns, and echoes of imperial shades still march patrols on vanished streets. To walk the Valewood is to risk being claimed by it: as guest, prey, heir---or trespasser.

\vspace{2mm}
\textbf{What you notice first:} The smell of wet moss and old stone. The way the light filters through leaves like stained glass. The silence that falls when you speak a name too loudly.

\textbf{The one rule that kills newcomers:} Never eat the fruit of a tree that has a face. The forest does not forget its debtors, and the first bite is always the price of admission.
\end{tcolorbox}

\subsection*{Regional Mood: Leaves, Oaths, and Echoes}

Here, the past is not past---it phases in and out like moonlight on water. The Lethai‑ar who remain are not conquerors but custodians, tending to laws written before the first human kingdom rose. The fae courts of Hazel and Alder play their endless games of courtesy and obligation, and the beast‑kin speak in parliaments where every vote is a scent. Imperial shades march through walls that no longer exist, and the star‑roads hum with geometries that twist the unwary into yesterday. The Valewood does not forgive arrogance. It expects humility, patience, and the willingness to trade a piece of yourself for safe passage.

\subsection*{Prominent NPCs of the Valewood}
\label{sec:valewood-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Queen Amalthea & Hazel Queen, sovereign of the Hazel Court & Preserve the old pacts and keep the empire from waking & She has never spoken a lie. She has never needed to. \\
\hline
King Thorn & Alder King, ruler of the Twilit Court & Collect debts, old and new, and expand the forest's reach & His crown is a noose of thorns. He wears it proudly. \\
\hline
Pathweaver Sila & Lethai‑ar guide (commoner exemplar) & Knot safe routes and keep the star‑roads from bleeding & Her cord has eight knots; the ninth is the one she never ties. \\
\hline
Bramble & Fox‑headed courier of the beast‑kin & Deliver messages that cannot be written & The price is always a secret, and the secret is always bitter. \\
\hline
The Huntsman & Antlered hunter of the Moonlit Ride & Claim those who break the forest's laws & His face is a stag's skull; his voice is the wind in the leaves. \\
\hline
Shardwright Kael & Artisan who repairs the star‑roads & Wake the ancient ways before they fade forever & Her chisel is a tooth from the first dragon. She stole it. \\
\hline
\end{longtable}

\subsection*{Names of the Valewood}
\label{sec:valewood-names}

Inspired by the old Lethai language and the fae courts: flowing vowels, soft consonants, and names that feel like they belong to a forgotten song.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Caelith & Amalthea & Thornweave & \textit{the Green} \\
\hline
Eryndor & Sila & Starfoot & \textit{Oath‑Woven} \\
\hline
Kael & Liriel & Hawthorn & \textit{the Unseen} \\
\hline
Oren & Vesper & Silverbough & \textit{Hollow‑Tongue} \\
\hline
Riven & Fenn & Mossglen & \textit{the Last} \\
\hline
Thorn & Myrin & Gloaming & \textit{Ash‑Finger} \\
\hline
\end{longtable}

\paragraph*{(Phasing Ruin/Star‑Road/Living Stone)} Star‑road shard with pale flagstones that hum when trod in sequence; rooted amphitheatre with moss‑seats remembering speeches.

\section*{Spades --- Places}
\label{sec:valewood-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Star-Road Shard} --- Pale flagstones hum when trod in sequence; a path between worlds. `[MAGIC]`
  \hfill \textit{The sequence is always incomplete. The missing step is the one you hesitate to take.}
\item \textbf{Rooted Amphitheatre} --- Moss‑seats remember speeches not yet given; a stage for fate. `[SOCIAL]`
  \hfill \textit{Sit in the high seat, and the roots will ask you a question. Answer wrong, and you may never stand.}
\item \textbf{Moon-Cistern} --- Reflecting sky that isn't tonight's; water as a mirror of truth. `[MAGIC]` `[TRUTH]`
  \hfill \textit{The reflection shows what you lost, not what you carry. Look away before it names you.}
\item \textbf{Glyphed Bridge} --- Across a dry stream; glyphs glow when lies pass; truth as a toll. `[TRUTH]` `[TRAVEL]`
  \hfill \textit{The bridge has no railings. Liars fall---into the stream, or into memory.}
\item \textbf{Glassleaf Gallery} --- Boughs grown into panes, dust like constellations; beauty as a barrier. `[MAGIC]`
  \hfill \textit{The dust is crushed star‑road. Breathe it, and you will dream of the empire's fall.}
\item \textbf{Hollow Aqueduct} --- Water runs uphill if the song is right; magic in the melody. `[MAGIC]`
  \hfill \textit{The song has eight verses. The ninth is the silence that follows.}
\item \textbf{Calendar Grove} --- Standing‑trees aligned to solstices; one trunk is stone; time as architecture. `[RITUAL]`
  \hfill \textit{The stone tree bleeds sap once a year. Collect it, and the forest owes you a season.}
\item \textbf{Unfound Arcade} --- Arches appear at dusk, vanish at dawn; reality as a schedule. `[MYSTERY]`
  \hfill \textit{The arches have names---Carved on the inside. You must walk through to read them.}
\item \textbf{Breathing City} --- Vine‑choked streets shift on the hour; a ruin as a living thing. `[DUNGEON]`
  \hfill \textit{The vines pulse with a slow heartbeat. The city is sleeping, not dead.}
\item[J] \textbf{Amber Ziggurat} --- Caught mid‑construction for a thousand years; time as an artifact. `[LEGEND]`
  \hfill \textit{The workers are statues of amber. Their eyes follow you.}
\item[Q] \textbf{Ivory Observatory} --- Brass orrery tracks two lost moons; the sky as memory. `[MAGIC]`
  \hfill \textit{The moons are not in any sky you know. One of them is yours.}
\item[K] \textbf{Throne-Bower} --- Living alder and argent wire; the seat of forgotten power. `[AUTHORITY]`
  \hfill \textit{Sit here, and the forest will crown you---or crush you. You won't know which until the leaves settle.}
\item[A] \textbf{Valeheart Spire} --- The empire's axle; stairs that end where they began. `[LEGEND]`
  \hfill \textit{The spire has no top. It has only a center. Finding it is the journey.}
\end{enumerate}

\paragraph*{(Lethai‑ar/Fae/Beast‑kin)} Pathweaver who knots safe routes into cord; fox‑headed courier with a promise braided into its tail.

\section*{Hearts --- People \& Factions}
\label{sec:valewood-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Pathweaver} --- Lethai‑ar guide who knots safe routes into cord; navigation as community. `[SOCIAL]`
  \hfill \textit{Her cord has eight knots. The ninth is the one she never ties.}
\item \textbf{Fox-Courier} --- Promise braided into its tail; messages as living things. `[SOCIAL]`
  \hfill \textit{It will deliver anything---for a price. The price is always a secret.}
\item \textbf{Owl-Sister} --- Pupils show yesterday; her hands show yours; sight as prophecy. `[MAGIC]`
  \hfill \textit{She never blinks. If she does, the future has changed.}
\item \textbf{Antler-Hunter} --- Sworn to keep the cities sleeping; a guardian of rest. `[COMBAT]`
  \hfill \textit{His spear is a sharpened root. It grows back when broken.}
\item \textbf{Moss-Scribe} --- Writes deeds in lichen, reads stone like vellum; law as growth. `[LAW]`
  \hfill \textit{Her ink is rainwater and regret. The words fade when the promise is kept.}
\item \textbf{Lark-Keeper} --- Caged bird sings warnings strangers can't hear; danger as song. `[OMEN]`
  \hfill \textit{The lark's song is a map. Learn it, and you can walk anywhere---once.}
\item \textbf{Green Neighbor} --- Fae courtier wearing a borrowed smile and boots; hospitality as a trap. `[FAE]`
  \hfill \textit{Never accept a gift from a Green Neighbor. It's never a gift.}
\item \textbf{Warden-Coterie} --- Beast‑kin (badger, stag, lynx); one law: no digging deep. `[COMMUNITY]`
  \hfill \textit{They speak in growls and gestures. The lynx is the one who remembers faces.}
\item \textbf{Echo-Legionary} --- Imperial shade bound to a patrol that never ends; duty as a curse. `[DEATH]`
  \hfill \textit{Its sword passes through stone but not through flesh. Run, don't fight.}
\item[J] \textbf{Shardwright} --- Ancient artisan who wakes star‑road with a chisel; craft as resurrection. `[CRAFT]`
  \hfill \textit{Her chisel is a tooth from the first dragon. She stole it.}
\item[Q] \textbf{Hazel Queen} --- Lethai‑ar sovereign in leaf and law; wisdom as authority. `[AUTHORITY]`
  \hfill \textit{She has never spoken a lie. She has never needed to.}
\item[K] \textbf{Alder King} --- Twilit Court sovereign in shadow and debt; power as obligation. `[AUTHORITY]`
  \hfill \textit{His crown is a noose of thorns. He wears it proudly.}
\item[A] \textbf{The Huntsman} --- Antlers of moonlight; shortcuts owe him; the path as price. `[DEATH]` `[TRAVEL]`
  \hfill \textit{His face is a stag's skull. His voice is the wind in the leaves.}
\end{enumerate}

\paragraph*{(Glamour/Ward‑Trap/Imperial Residue)} Sweet wind nudges landmarks hedge‑width---enough to matter; path reverses---footprints vote to go elsewhere.

\section*{Clubs --- Complications/Threats}
\label{sec:valewood-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Sweet Wind} --- Nudges landmarks hedge‑width---enough to matter; navigation as deception. `[WEATHER]` `[TRICK]`
  \hfill \textit{The wind smells of honey and lies. Do not trust your nose.}
\item \textbf{Path Reversal} --- Footprints vote to go elsewhere; direction as democracy. `[MAGIC]`
  \hfill \textit{Your tracks point the way you came---even when you haven't turned.}
\item \textbf{Ward-Trap} --- Ancient security treats you as yesterday's invader; the past as prison. `[TRAP]`
  \hfill \textit{The glyphs were set by a dead queen. She still enforces her grudges.}
\item \textbf{Oath-Magnet} --- Spoken promises stick and reshape plans; word as binding magic. `[OATH]`
  \hfill \textit{Every ``I will'' becomes a knot. Every ``I won't'' becomes a chain.}
  \begin{tcolorbox}[colback=red!3,title=Oath-Magnet Mechanics]
    When a character makes a promise in the Valewood, the GM creates a \textbf{Promise Clock [4]}. Each time the character acts against that promise (or fails to act in its favour), tick the clock. When it fills, the forest extracts a toll: 1 Fatigue, a lost memory (erase a minor fact or face), or a minor geas (the GM chooses). Breaking a promise intentionally (not through failure) instantly ticks the clock twice.
  \end{tcolorbox}
\item \textbf{Geas Catch} --- A careless word triggers a task that names itself; speech as summoning. `[OATH]`
  \hfill \textit{You said ``if only.'' Now the forest has heard you.}
\item \textbf{City Phase} --- Streets rotate; the exit becomes a balcony; space as puzzle. `[DUNGEON]`
  \hfill \textit{The city breathes. You are in its lungs.}
\item \textbf{Name-Theft} --- Something learns your true name's edges; the tests begin. `[DEATH]` `[FEAR]`
  \hfill \textit{First, you forget a minor debt. Then a friend's face. Then your own.}
  \begin{tcolorbox}[colback=red!3,title=Name-Theft Mechanics]
    When \textbf{Name-Theft} occurs, the GM secretly records one piece of the target's identity (a memory, a bond, a talent name). The target suffers –1 die on any roll that would rely on that piece until it is restored by a quest or a bargain with the thief (often a fae or shade).
  \end{tcolorbox}
\item \textbf{Ring Claim} --- A circle names you host; feeding guests becomes law. `[SOCIAL]` `[COMMUNITY]`
  \hfill \textit{The mushrooms are listening. They will remember what you serve.}
\item \textbf{Mirror Rain} --- Leaf‑drips replay choices you didn't make; the past as echo. `[MAGIC]`
  \hfill \textit{The rain shows you a path you refused. Now it offers it again---with interest.}
\item[J] \textbf{Redcaps Abroad} --- Hats wet, boots quick, courtesies thin; violence as hospitality. `[COMBAT]` `[FAE]`
  \hfill \textit{They will ask for a kiss or a coin. Refuse, and they take both.}
\item[Q] \textbf{Court Tithe} --- Truth, song, or memory you'd miss; the price as sacrifice. `[SOCIAL]` `[OATH]`
  \hfill \textit{The courtier opens its hand. What you place there is never returned.}
\item[K] \textbf{Muster of Boughs} --- Green banners rise; travel becomes trespass. `[WAR]` `[FEAR]`
  \hfill \textit{The boughs bear old sigils. The war they remember ended a thousand years ago.}
\item[A] \textbf{Empire Wakes} --- A district aligns; doors open, guardians open farther. `[APOCALYPSE]`
  \hfill \textit{The shades have orders. You are the target.}
\end{enumerate}

\paragraph*{(Charm/Key/Truce‑Bough)} Way‑cord knot points to the true next turn; dew‑mirror shows a thing as it is, not as it is sung.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:valewood-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Way-Cord} --- A knot points to the true next turn (once); navigation as a gift. `[TRAVEL]`
  \hfill \textit{The knot is a breath tied in silk. Untie it to find your way---or to lose it forever.}
\item \textbf{Dew-Mirror} --- See a thing as it is, not as it is sung, for one scene; truth as a tool. `[TRUTH]`
  \hfill \textit{The mirror is a drop of water on a leaf. It dries when the lie dies.}
\item \textbf{Hazel Token} --- Lawful crossing of a warded hedge without snag; permission as paper. `[TRAVEL]`
  \hfill \textit{The token is a hazelnut. Crack it open; the kernel is your password.}
\item \textbf{Honey-Right} --- Eat and speak safely at any bee‑stone ring; hospitality as law. `[SOCIAL]`
  \hfill \textit{The honey is golden. Taste it, and the bees will know you as a guest---once.}
\item \textbf{Name-Bead} --- A kept promise warms; a broken one chills and glows; an oath as a magic item. `[OATH]`
  \hfill \textit{The bead is warm in your palm. If it cools, you have forgotten someone.}
\item \textbf{Wind-Veil} --- Mute a sweet wind's lies for a short walk; truth as protection. `[WEATHER]`
  \hfill \textit{The veil is a scrap of silk. Breathe through it, and the wind cannot speak.}
\item \textbf{City-Shard} --- Wake one gate or stair in a phasing ruin; a key as a catalyst. `[MAGIC]`
  \hfill \textit{The shard is a piece of a star‑road. Hold it to the lock, and the ruin answers.}
\item \textbf{Truce-Bough} --- Hunt, cut, or camp without offense in a named copse; peace as a branch. `[SOCIAL]`
  \hfill \textit{The bough is green. Plant it in the ground, and the forest will not harm you---or help you.}
\item \textbf{Oathsap} --- Seal a pact that even echoes respect; a promise as a binding force. `[OATH]`
  \hfill \textit{The sap is warm. Drink it, and your words become stone---for good or ill.}
\item[J] \textbf{Shardwright's Favor} --- Repair or light a length of star‑road; craft as power. `[CRAFT]`
  \hfill \textit{The favor is a chip of amber. The Shardwright will come when you break it.}
\item[Q] \textbf{Hazel Audience} --- One whisper exits as policy; word as law. `[AUTHORITY]`
  \hfill \textit{The audience is a leaf. Speak into its spine, and the Hazel Queen will listen---once.}
\item[K] \textbf{Alder Writ} --- Levies defer, wardens guide, scribes stop arguing; authority as writ. `[AUTHORITY]`
  \hfill \textit{The writ is a thorn. Carry it in your boot, and the Alder King's debtors will make way.}
\item[A] \textbf{Valeheart Clause} --- A temporary exception to any local rule, spoken in leaves. `[LEGEND]`
  \hfill \textit{The clause is a rustle. The forest will pretend it didn't hear---but it will remember.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:valewood-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (charms/keys/passes) that change position rather than call for a roll.
\item Empire Echo: when a J/Q/K is the highest rank in your seed, add one visible relic‑logic (floating stairs, singing locks, patrol shades).
\item If any A appears, echo \textbf{leaf and light} motifs---paths that remember, leaves that whisper, light that moves like water.
\end{itemize}

\subsection*{Additional Features}
\begin{itemize}
\item \textbf{Forest Claims:} Any oath, bargain, or gift exchanged binds with more weight. Once per session, a spoken promise becomes magically enforceable: breaking it creates 1d4 SB and echoes through scenes.
\item \textbf{Ruins Awaken:} If you linger at an ancient site, roll or draw to see which mechanism stirs. May grant guidance (lit star‑road, a memory shown) or danger (patrol shades, collapsing glyphs).
\item \textbf{Court Interference:} Encounters with fae or beast‑kin are never neutral. Once per session, the GM may convert a Diamond into a demand for hospitality, tribute, or alliance---shifting the expedition's tone.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Unseen Court} — The Witness. *A parliament of silent roots that remembers every broken promise.*
    \item \textbf{The Root-King} — The Alder King (local aspect). *Debt that grows like a vine, choked by its own leaves.*
    \item \textbf{The Star-Weaver} — The Ninth. *The missing thread in every star‑road, the silence after the song.*
    \item \textbf{The Green Saint} — Oath of Flame \& Light (syncretic). *A martyr who fed her own name to the Hollow to seal a breach.*
  \end{itemize}
\end{itemize}

\subsection*{Forest Mood (d6 per scene)}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title=Forest Mood Table]
\begin{enumerate}
\item \textbf{Hazel's Favour} – Paths open willingly. +1 die to navigation rolls this scene.
\item \textbf{Alder's Hunger} – Debts are called. Someone in the scene must offer a memory or service to proceed.
\item \textbf{Empire Wakes} – A patrol of imperial shades marches through. Avoid or pay a toll (1 Fatigue or a secret).
\item \textbf{Beast-kin Moot} – Animals gather and watch. They will not attack unless provoked, but they remember faces.
\item \textbf{Sweet Wind} – All navigation rolls suffer –1 die; landmarks shift subtly.
\item \textbf{Green Gate Breath} – A threshold appears. Crossing it leads to a different part of the forest (or a different time).
\end{enumerate}
\end{tcolorbox}

\subsection*{Star-Road Discovery (d6)}
When the party encounters a star-road shard or attempts to wake a segment, roll on this table for an additional feature:
\begin{enumerate}
\item \textbf{Singing Stones} – The flagstones hum a tune. Anyone who sings along gains +1 die on the next roll.
\item \textbf{Echo of an Emperor} – A ghostly figure appears and asks a riddle. A correct answer grants a \textbf{City-Shard}.
\item \textbf{Fractured Timeline} – The road shows a glimpse of the future (1d6 scenes ahead). The GM must answer one yes/no question truthfully.
\item \textbf{Ward-Trap} – The segment is booby-trapped. Trigger a \textbf{Ward-Trap} complication (see Clubs).
\item \textbf{Shardwright's Cache} – Hidden nearby is a \textbf{Shardwright's Favor} (Diamond) or a rare component (e.g., amber shard).
\item \textbf{The Ninth Step} – The sequence has a hidden ninth stone. Stepping on it teleports the party to a \textbf{Valeheart Spire} or a court audience.
\end{enumerate}

\subsection*{Beast-kin Traits (for NPCs or PC ancestry)}
Beast-kin are not a single species but a loose alliance of animal‑aspected beings. Use these traits for quick NPCs or optional player characters.
\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title=Beast-kin Traits]
\begin{itemize}
  \item \textbf{Scent Memory:} +1 die to track a creature you have smelled before.
  \item \textbf{Parliament Oath:} Once per session, you may invoke the beast-kin moot to ask a local animal one simple question (the GM answers truthfully).
  \item \textbf{Claw or Fang:} Unarmed strikes deal Harm 1 and count as Light weapons.
  \item \textbf{Optional:} Choose a lineage:
  \begin{itemize}
    \item \textit{Stag:} +1 die to resist being moved or knocked prone.
    \item \textit{Lynx:} +1 die to stealth in forest terrain.
    \item \textit{Badger:} You may burrow through loose earth at half speed.
  \end{itemize}
\end{itemize}
\end{tcolorbox}

\subsection*{Lore Echoes (Cross‑Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Genesis Vats (The Clockwork Monad):} Deep beneath the Amber Ziggurat lie the empire's failed laboratories. Here, the Lethai ancients combined elf and beast, hoping to create the perfect guardian. The vats still hum, and the beast‑kin who survived do not speak of what they saw. The Shardwrights say the Monad's pressure is still optimising the old designs, and that one day the vats will open again.
  \item \textbf{The Parliament of Claw and Fur (The Savage Heart):} The beast‑kin are not savages. They hold a moot once every season, deep in a clearing that no star‑road touches. Their laws are carved in scent and scar, and their judgments are final. The Savage Heart speaks through their eldest, but it does not command—it suggests. Outsiders who attend the moot must bring a true name of a foe and leave it on the speaking stone. Those who lie are given fur and four legs, and are hunted at the next moon.
  \item \textbf{Saint Oren of the Unfinished Star (The Ninth):} A Lethai‑ar astronomer who calculated the exact angle at which the empire's star‑roads would collapse. He was erased from the imperial records, but his ghost still walks the Ivory Observatory, tracing equations in starlight. Those who find him may ask one question about the forest's hidden geometry. The answer is always a single number, and that number is a debt.
  \item \textbf{The Grey Pilgrims (The Traveler):} A wandering order of Lethai‑ar who maintain the old way‑cords. They wear grey cloaks woven from mist and speak only in riddles. It is said that if you give a Grey Pilgrim a lock of your hair, they will knot it into your way‑cord, and you will never be lost in the Valewood again—but you will also never find a place you truly wish to leave.
  \item \textbf{The Daughters of the Cord (The Bitter Root):} A renegade Sister hides among the moss‑scribes of the Rooted Amphitheatre, trying to untie her own cord from the Hazel Queen's ledger. She claims the queen's hospitality is a cage, and that the forest's memory is a trap. The Hazel Queen has not corrected her; she only smiles, and the roots grow tighter each night.
  \item \textbf{The Ninth:} In any Valewood scene, the ninth leaf to fall from a tree will land pointing to a hidden threshold. The wise catch it---or pretend they didn't see. Those who follow the leaf's direction find a place that was not there a moment ago, and something old waits inside.
\end{itemize}

\subsection*{Valewood Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Roots:} Crumble a crust at the base of a standing stone. Ignore the first \emph{Sweet Wind} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Cord:} Tie a single knot in a way‑cord. Cancel \emph{Oath-Magnet} or take –1 die on your next promise (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Cistern:} Whisper a dead enemy's name into the Moon‑Cistern. Gain +1 Effect when evading fae this scene, but mark \emph{Obligation} to the Alder King. `[SUPERSTITION]`
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In the Valewood, power flows through ancient magic, forgotten oaths, and the delicate balance between guest and host. The Lethai‑ar maintain influence through their knowledge of paths and ruins, while the fae courts wield power through hospitality and geas. Beast‑kin serve as guardians of the deeper mysteries, and the echoes of the empire still patrol their eternal rounds, enforcing laws written in starlight and stone.

\textbf{For the GM:}  
Patronage in the Valewood revolves around hospitality, knowledge, and the careful navigation of ancient laws. Rewards often take the form of charms, keys, or safe passage that can be leveraged into greater influence. To emphasize this:
\begin{itemize}
\item Tie rewards to visible symbols (tokens, mirrors, cords) that can be challenged, stolen, or voided.
\item Let rival courts issue conflicting protections, forcing players to choose whose favour matters more.
\item Use the ruins, groves, and star‑roads as arenas for social contests, where knowledge of ancient customs determines success.
\end{itemize}
In the Valewood, your word is your world, and your world determines whether you walk as a guest or a ghost.
\end{tcolorbox}

% Index entries
\index{Valewood|see{Empire Under Leaves}}
\index{Theme generator}
\index{Empire Under Leaves generator}
\index{Places!Valewood}
\index{People!Valewood}
\index{Complications!Valewood}
\index{Rewards!Valewood}
\index{Hazel Queen}
\index{Alder King}
\index{The Huntsman}
\index{Star-Road}
\index{Rooted Amphitheatre}
\index{Glyphed Bridge}
\index{Moon-Cistern}
\index{Breathing City}
\index{Valeheart Spire}
\index{Pathweaver}
\index{Owl-Sister}
\index{Moss-Scribe}
\index{Echo-Legionary}
\index{Shardwright}
\index{Way-Cord}
\index{Dew-Mirror}
\index{Valeheart Clause}
\index{Beast-kin}
\index{Genesis Vats}
\index{Grey Pilgrims}
\index{Forest Mood}
\index{Star-Road Discovery}
\index{Local Patrons!Valewood}

\section*{Thematic SB Spend Table}
\label{sec:valewood-sb}

\subsection*{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Star-Road Flicker} – A nearby star-road shard emits a pulse of light. Your position becomes visible to imperial shades; the next stealth roll suffers –1 die.
\item \textbf{Hazel's Displeasure} – A fae courtier notes your lack of courtesy. The next social roll with any fey is at –1 Position.
\item \textbf{Memory Echo} – A ghostly whisper repeats a secret you told years ago. The GM gains 1 SB to spend on a future complication involving that secret.
\item \textbf{Leaf Whisper} – The forest carries your words to an unintended listener. A minor faction becomes aware of your presence (mark Exposure +1).
\item \textbf{Vine Snag} – Roots clutch at your feet. Delay 1 segment as you free yourself.
\end{itemize}

\subsection*{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Path Reversal (Intensified)} – The trail you were following now leads back to where you started. Navigation DV +2 until you find a new approach.
\item \textbf{Oath-Magnet Surge} – A careless promise binds tighter than intended. The GM creates a Promise Clock [4] immediately; tick it once.
\item \textbf{Imperial Patrol Shifts} – A squad of Echo-Legionaries phases into your corridor. They are not hostile unless you carry imperial relics or speak the old tongue.
\item \textbf{Geas Catch (Minor)} – The forest interprets a casual phrase as a vow. You must perform a small task (e.g., “never eat honey again”) or mark 1 Fatigue.
\item \textbf{Fey Tithe Demanded} – A Green Neighbor appears and demands a gift. Refuse and lose 1 Favour with the Hazel Court; comply and the GM gains 1 SB.
\end{itemize}

\subsection*{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Name-Theft Attempt} – A shade or fae learns a fragment of your true name. Mark \emph{Name Leverage} [1] (the first segment). The GM may invoke it once.
\item \textbf{Ring Claim (Forced Hospitality)} – The mushrooms close in. You must host a group of fae for one scene; failure to entertain them triggers a Court Tithe (memory or story).
\item \textbf{City Phase Lockdown} – A phasing ruin shifts while you are inside. The exit vanishes. You must solve a puzzle or survive a Ward-Trap to escape.
\item \textbf{Empire Echo Intensifies} – A district of the Breathing City aligns. Imperial shades receive new orders: capture, not kill. They gain +1 die to grapple and detain.
\item \textbf{Alder's Debt Called} – The Alder King sends a collector. You owe a favour (from an earlier scene). Pay with a story, a memory, or a service.
\end{itemize}

\subsection*{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Star-Road Awakening} – A forgotten star-road segment activates. The party is pulled into a Between‑Places shortcut – but a Shardwright's Favour or a True Name is required to exit.
\item \textbf{The Huntsman's Horn} – The Moonlit Ride begins. All fae gain +1 Position against you until you offer a truce (Truce-Bough) or defeat the Huntsman in a riddle contest.
\item \textbf{Empire Wakes (Full)} – A sealed district of the Breathing City opens. Ancient guardians patrol every corridor, and the Hollow's attention is drawn (start Hollow Attention [4]).
\item \textbf{Hazel Queen's Verdict} – The Hazel Queen herself appears. She passes judgment on a past action (real or imagined). You may plead your case; failure costs a cherished memory.
\item \textbf{The Ninth Leaf Falls} – A leaf points to a threshold that was not there before. Crossing it leads to the Hollow's Court. The Oath-Keeper awaits.
\end{itemize}

\subsection*{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Valewood (Star-Roads):} A shard hums out of phase; stepping stones rearrange into a new sequence; a forgotten geometry twists your path.
\item \textbf{Valewood (Fae Law):} A courtier demands a toll in courtesy; a gift is refused and becomes a curse; a promise is heard by leaves and roots.
\item \textbf{Valewood (Imperial Echoes):} A patrol of shades marches through solid rock; a sealed door groans and cracks; the ghost of a Lethai‑ar scribe appears and offers a deal.
\end{itemize}

\subsection*{Boss Seeds (three‑phase foes)}
\begin{itemize}
\item \textbf{The Huntsman of Shortcuts (antlers of moonlight)} --- \emph{Phases}: Owe the Cut $\rightarrow$ Step Across $\rightarrow$ Tally the Tread. \emph{Cracks}: present a \textbf{Truce-Bough} to halt pursuit; a \textbf{Way-Cord} cancels \emph{Step Across} once.
\item \textbf{Redcap Captain Rill (hat never dries)} --- \emph{Phases}: Hospitality Turned $\rightarrow$ Knee the Truth $\rightarrow$ Boot‑Hook. \emph{Cracks}: \textbf{Honey-Right} reverses one phase; \textbf{Dew-Mirror} exposes his ``host right.''
\item \textbf{Echo Centurion of the XIII Star-Road (bronze mask)} --- \emph{Phases}: Reform the Line $\rightarrow$ Juridical Recall $\rightarrow$ Standard Rekindled. \emph{Cracks}: a \textbf{City-Shard} lets you counter‑order once; a \textbf{Moss-Scribe} ruling cancels \emph{Juridical Recall}.
\item \textbf{Shardwright Unmended (glassy chisels)} --- \emph{Phases}: Cut the Clock $\rightarrow$ Wrong Stone $\rightarrow$ Wake the Segment. \emph{Cracks}: offer \textbf{Oathsap} to bind a repair oath; \textbf{Shardwright's Favor} flips \emph{Wrong Stone}.
\item \textbf{The Hazel Advocate (mask of living bark)} --- \emph{Phases}: Measure the Word $\rightarrow$ Prune the Lie $\rightarrow$ Court Silence. \emph{Cracks}: \textbf{Hazel Token} grants speaking rights; \textbf{Oathsap} lets you safely retract one bound clause.
\end{itemize}

\subsection*{Legendary Sites \& Trials}
\begin{itemize}
\item \textbf{Glyphed Bridge Ordeal} --- Speak a hard truth to pass dry; lie, and the bridge demands a memory tithe (mark \emph{Name Leverage} +1, gain \emph{Bridge's Favor} only if confessed).
\item \textbf{Moon-Cistern Vigil} --- Watch someone else's sky until the owl calls; success grants a single \textbf{Dew-Mirror} charge; failure spawns \textbf{Mirror Rain}.
\item \textbf{Valeheart Spiral} --- Climb stairs that return to start; map with verse, gift, count, listen (all four once) to unlock a \emph{Between‑Places} shortcut.
\end{itemize}

% ====== VALEWOOD: PLUG-IN EXTENSIONS ======
\subsection*{Extensions (Plug-in)}
\label{sec:valewood-extensions}
\begin{itemize}
  \item \textbf{Name Economy (true‑name pressure):}
  \begin{itemize}
    \item Each PC has \emph{Name Facets}: \textbf{Called‑Name} (common), \textbf{Given‑Name} (trusted), \textbf{True‑Edge} (dangerous).
    \item When a foe or court learns a deeper facet, mark \emph{Name Leverage} [3]. At 3, they may compel one precise action or forbid one path unless countered by a \Diamond{} or \textbf{Oathsap}.
    \item You may trade a shallower facet to erase 1 tick from \emph{Name Leverage} but gain \emph{Rumor Tail} condition until scene end.
  \end{itemize}

  \item \textbf{Hospitality Ledger (guest/host law):}
  \begin{itemize}
    \item When \emph{Honey‑Right} or \emph{Truce‑Bough} is invoked, set a 3‑step \emph{Table Owed} clock for the host and a 3‑step \emph{Courtesy Owed} for the guest.
    \item Ticking your clock grants +1d to social tests with that party; overpaying (4+) converts 1 tick into a portable \emph{Favor Twig} (spend for +1 Position later).
    \item Breaking either clock spawns \textbf{Ring Claim} or \textbf{Oath‑Magnet} at the keeper's option.
  \end{itemize}

  \item \textbf{Path‑Logic Mini‑game (finding the true way):}
  \begin{itemize}
    \item Two opposed clocks per journey: \emph{True Way} vs \emph{Forest Has You} (size by highest rank in the seed).
    \item Each leg, pick an \emph{Approach}: \textbf{Rhyme} (speak a pattern), \textbf{Count} (steps/time), \textbf{Gift} (leave a token), \textbf{Listen} (ambient omen).
    \item Edge: Rhyme\(\rightarrow\)Gift (courtesies bind), Gift\(\rightarrow\)Count (the land answers), Count\(\rightarrow\)Listen (timing frames the omen), Listen\(\rightarrow\)Rhyme (echo feeds verse).
    \item Spend a \textbf{Way-Cord} to flip a failed leg; spend a \textbf{Wind-Veil} to ignore \textbf{Sweet Wind} once.
  \end{itemize}

  \item \textbf{Truth Mechanics (seeing through glamour):}
  \begin{itemize}
    \item \textbf{Dew-Mirror} reveals one \emph{Unkind Fact} in the scene; naming it aloud gives +1 Effect but ticks \textbf{Court Tithe} +1.
    \item \textbf{Glyphed Bridge} judges the next lie uttered; if you pass speaking truth, gain \emph{Bridge's Favor} (+1 Position on crossings hereafter). If you lie, mark \textbf{Name‑Theft} +1.
  \end{itemize}

  \item \textbf{Star‑Road Procedure (phasing traversal):}
  \begin{itemize}
    \item To wake a segment: meet two of three---\emph{Proper Footfall} (count/pattern), \emph{Shard Key} (City‑Shard, Shardwright help), \emph{Witness} (owl, echo‑legionary, or courtier).
    \item Success grants \emph{Between‑Places}: one safe shortcut and one \emph{Echo Fee} (memory, song, or small timeline bruise chosen by the GM).
    \item On a miss, spawn \textbf{Ward‑Trap} or \textbf{City Phase}.
  \end{itemize}

  \item \textbf{Court Etiquette (Hazel vs Alder):}
  \begin{itemize}
    \item \textbf{Hazel Audience}: pays in \emph{precision}---exact words, measured gifts. +1d when you cite prior custom or present a \textbf{Moss‑Scribe} ruling.
    \item \textbf{Alder Writ}: pays in \emph{obligation}---future tasks, debts. +1 Effect when you promise service; tick \textbf{Oath‑Magnet} +1.
  \end{itemize}

  \item \textbf{Start‑of‑Scene Omens (1d6):}
  \begin{enumerate}
    \item \emph{Leaf‑Whisper}: first Listen/Study gets +1d; first boast draws \textbf{Geas Catch}.
    \item \emph{Antler Shadow}: the Huntsman watches---shortcuts improve; hiding worsens.
    \item \emph{Mirror Drip}: \textbf{Dew-Mirror} free use but \textbf{Court Tithe} starts at 1.
    \item \emph{Root Push}: terrain shifts in your favor once; buildings dislike you (locks complicate).
    \item \emph{Bird Parliament}: a rumor arrives true; one plan becomes public knowledge.
    \item \emph{Star‑Tingle}: star‑road easier (+1d to wake) but every \Diamond{} hereafter risks \textbf{Ward‑Trap} on 1–2.
  \end{enumerate}

  \item \textbf{Plug‑in SB Conversions (Valewood Flavor):}
  \begin{itemize}
    \item 1 SB \(\rightarrow\) \emph{Sweet Wind Shift}: reposition an approach vector; –1d unless you adjust method (Rhyme/Count/Gift/Listen).
    \item 2 SB \(\rightarrow\) \emph{Obligate Courtesy}: immediate small gift or truth required; refuse and tick \emph{Name Leverage} +1.
    \item 3+ SB \(\rightarrow\) \emph{Empire Echo}: add a patrolling \textbf{Echo‑Legionary} cohort or activate a dormant mechanism that reframes the scene.
  \end{itemize}

  \item \textbf{Reputation Echoes:}
  \begin{itemize}
    \item \emph{Guest‑Faithful}: once/session, downgrade \textbf{Ring Claim} to simple hospitality; the next refused invitation spawns \textbf{Oath‑Magnet}.
    \item \emph{Path‑Favored}: +1 Position on \emph{Path‑Logic} legs; any shortcut taken owes a \emph{token or tale}.
    \item \emph{Truth‑Bearer}: +1d with the Hazel court when presenting facts; –1d with the Alder court when declining obligation.
  \end{itemize}

  \item \textbf{Quick Cross‑Region Hooks:}
  \begin{itemize}
    \item \emph{Kahfagia Lanes \(\leftrightarrow\) Star‑Roads}: a \textbf{Mirror Warrant} can stand in for a \textbf{Shard Key} once.
    \item \emph{Mistlands Bells \(\leftrightarrow\) Path‑Logic}: paying \textbf{Ward‑Salt} counts as the \emph{Gift} approach with +1d.
    \item \emph{Aelinnel Math \(\leftrightarrow\) Proper Footfall}: \emph{Counting Etiquette} grants +1 Position on the \textbf{Star‑Road Procedure}.
  \end{itemize}
\end{itemize}

\subsection*{Artifacts of the Valewood}
\label{sec:valewood-artifacts}

The Valewood remembers not only oaths and names but also the tools of those who came before. Some artifacts are cursed by the empire’s fall; others are blessed by the forest’s slow patience. Each artifact carries its own burden—a fragment of will that accumulates Obligation as it is used. When the Obligation fills, the artifact demands a price, just as a Patron does.

\subsubsection*{Artifact Obligation Rules}
\begin{itemize}
  \item An artifact has an \textbf{Obligation Capacity} equal to the wielder’s \textbf{Spirit} (or a fixed number, as noted). Mark Obligation each time a Rite is invoked.
  \item When Obligation exceeds Capacity, the wielder marks 1 Fatigue per extra segment (as per standard Obligation overflow).
  \item When the artifact’s Obligation clock reaches double its Capacity, it \textbf{demands} a price (see each artifact’s “Demand”). Refusing the demand may cause the artifact to become \textit{Compromised} (unusable until a quest is completed) or to inflict a permanent curse.
  \item Obligation can be cleared through acts of service aligned with the artifact’s nature (e.g., protecting a grove, telling a forgotten truth, shedding your own blood).
  \item Bound by Obligation: An artifact cannot be permanently discarded or destroyed while you owe it Obligation. If you deliberately abandon it (or attempt to destroy it by mundane means), it reappears beside you at the next dawn, and you mark +1 Obligation. This penalty applies only once per artifact; subsequent attempts to abandon it cause the artifact to become Compromised – it still reappears, but its Rites cannot be invoked until you perform a minor act of devotion (GM’s choice). An artifact can be legitimately transferred to another willing creature, clearing your Obligation; the new owner inherits any remaining Obligation. An artifact can be permanently destroyed only by a specific method (e.g., a star‑road shard, a dragon’s tooth, or the Hazel Queen’s permission), which typically requires a quest.
\end{itemize}

\begin{longtable}{|p{0.15\textwidth}|p{0.3\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{Artifact} & \textbf{Description} & \textbf{Rites (Abilities)} & \textbf{Demand (Obligation 2×Cap)} \\
\hline
\textit{Echo’s Lament} & A silver hand-bell with a cracked clapper. When rung, it repeats the last words spoken at the site of an imperial execution. & \textbf{Rite of the Last Echo (Low)} – Ask one question about a death that occurred within Near range; the bell answers with a whisper. Mark +1 Obligation. & The wielder must pronounce a true name of someone they have wronged, aloud and unasked, before the next full moon. \\
\hline
\textit{Thorn-Fang} & A dagger carved from a single hawthorn root, its edge weeping sap that smells of iron and honey. & \textbf{Rite of the Unhealing Prick (Low)} – On a successful melee attack, the wound bleeds for 1 additional Fatigue per scene (non‑stacking). Mark +1 Obligation. & The dagger demands a drop of the wielder’s own blood before each dawn; if refused, it turns on its owner (next attack targets self). \\
\hline
\textit{Glassleaf Lantern} & A lantern whose panes are made of petrified leaves. It burns without oil, casting light that reveals only things that are lost. & \textbf{Rite of the Unfound Path (Standard)} – Once per scene, the lantern illuminates a hidden threshold or a forgotten memory in the area. Gain +2 dice to Investigation or Navigation for that task. Mark +2 Obligation. & The lantern demands to be left at a crossroads at midnight. If not, it will guide the wielder into a dead end (next travel leg automatically fails). \\
\hline
\textit{Cord of Eight Knots} & A finger‑thick rope of braided grey wool, with eight tight knots. The ninth knot is missing. & \textbf{Rite of the Unbroken Promise (Low)} – Tie the cord around a willing creature’s wrist; until untied, that creature cannot willingly break a spoken oath. Mark +1 Obligation. & The cord demands a ninth knot, which must be tied in a strand of hair from someone who has broken a vow. Untying any existing knot costs 1 Fatigue. \\
\hline
\textit{Alder King’s Tooth} & A black, curling tooth set in a copper ring. When worn, it whispers the debts of those nearby. & \textbf{Rite of Debt-Sense (Low)} – Concentrate for a moment; learn the single greatest debt owed by a creature you can see (in vague terms: “a life,” “a stolen heirloom”). Mark +1 Obligation. & The tooth demands that you collect a minor debt (a few coins, a favour) from someone weaker than you before sunset. If you refuse, it begins listing your own debts aloud. \\
\hline
\textit{Shardwright’s Mallet} & A stone hammer with a handle of petrified wood. One side is smooth (to repair), the other rough (to break). & \textbf{Rite of the Broken Seal (Standard)} – Strike a phasing ruin or star‑road shard. Either temporarily stabilise it (open a safe passage) or shatter it (destroy the path, cause a Ward‑Trap). Mark +2 Obligation. & The mallet demands that you break something precious to you (a gift, a memento, a tool) and leave the pieces on a star‑road. \\
\hline
\textit{Mirror of Unfinished Faces} & A hand mirror with a silvered surface that shows not your reflection but a version of you who made a different choice. & \textbf{Rite of the Other Self (Low)} – Gaze into the mirror; gain +1 die on your next roll that involves a decision (the other self’s memory guides you). Mark +1 Obligation. & The mirror demands that you speak aloud the choice you regret most, to a stranger, within a day. If you refuse, the other self begins to appear in reflections, whispering doubts. \\
\hline
\textit{Whispering Horn} & A curved horn of amber, cracked along one side. When blown, it sounds like the roar of a forgotten beast. & \textbf{Rite of the Imperial Hunt (Standard)} – Once per session, blow the horn to summon a spectral pack of hounds that track a designated target for one scene (grants +1 die to tracking and pursuit). Mark +2 Obligation. & The horn demands a hunt – you must chase and catch something (even a mouse) within the hour, or the spirits will hunt you instead (suffer Harm 1 from ethereal bites). \\
\hline
\textit{Grave-Moss Cloak} & A heavy cloak woven from moss that grows only on imperial barrows. It is always damp and smells of rain. & \textbf{Rite of the Walking Dead (Low)} – Pull the hood up; for one scene, imperial shades treat you as one of their own (no attacks, you may pass patrols). Mark +1 Obligation per scene of use. & The cloak demands that you bury something that was never buried – a memory, a secret, a tooth. If you fail to do so before the next moon, it wraps itself around your throat in your sleep. \\
\hline
\textit{Hazel Queen’s Comb} & A wooden comb carved with thorned vines. Running it through your hair removes tangles and also removes recent memories. & \textbf{Rite of the Forgetful Stroke (Low)} – Stroke the comb through your hair; erase one recent short‑term memory (a face, a password, a few minutes of events). The memory is stored in the comb. Mark +1 Obligation. & The comb demands that you give a stored memory to a stranger without explanation. If you refuse, the comb begins to remove random memories each dawn (GM chooses). \\
\hline
\end{longtable}

\subsubsection*{Using Artifact Rites in Play}
\begin{itemize}
  \item Activating a Rite takes an action unless noted otherwise.
  \item Unlike Patron Rites, artifact Rites do not require a Codex or Familiar – only possession of the artifact.
  \item The GM may allow a wielder to clear artifact Obligation by performing a small act of devotion (e.g., for \textit{Thorn-Fang}: cut a thorn bush; for \textit{Glassleaf Lantern}: leave it in a moonbeam overnight).
  \item Artifacts can be stolen, lost, or destroyed. If an artifact is destroyed, any accumulated Obligation is cleared, but the destruction itself may trigger a final demand (GM’s discretion).
\end{itemize}

\index{Artifacts!Echo's Lament}
\index{Artifacts!Thorn-Fang}
\index{Artifacts!Glassleaf Lantern}
\index{Artifacts!Cord of Eight Knots}
\index{Artifacts!Alder King's Tooth}
\index{Artifacts!Shardwright's Mallet}
\index{Artifacts!Mirror of Unfinished Faces}
\index{Artifacts!Whispering Horn}
\index{Artifacts!Grave-Moss Cloak}
\index{Artifacts!Hazel Queen's Comb}
