% ============================================================
% Vhasia --- ``The Fractured Sun''
% Enhanced with dynastic marital intrigue from Province Wars era
% Revised to include Ninth Taboo manifestation and consistent formatting
% ============================================================
\section{Vhasia --- ``The Fractured Sun''}
\label{chap:vhasia}

\textbf{Key Demographics: } Humans
\textbf{Capital City: } Lence

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{The sun shattered at Lence. The pieces are still falling. Which shadow will catch you?}

\begin{quote}
  \textbf{Cardinal-Regent of the Broken Sun (Elite):}  
  ``The sun may be fractured, but our authority is absolute within our domains. Every coin minted here carries our blessing, and every law written bears our seal---the rest is just politics. The rest is also ash. Ask the Fenwoods. They made the ash, and they left us to choke on it. But remember: the first crack came not from their swords, but from the marriage beds where a Vhasian queen bore sons to two Viterran kings---and the sons who wondered whose blood truly ran in their veins. Chivalry? A thin veneer. Scratch it, and you find the iron beneath.''
\end{quote}

\begin{quote}
  \textbf{Road Warden (Commoner):}  
  ``I've served under three banners on this same road. The only constant is that someone always wants to collect tolls, oaths, the dead. The Fenwoods burned Lence ten years ago, marching from the Belworth like storm-clouds. Now every duke with a purse thinks he's king. The Queen-Mother weaves in the west, Valdais sharpens his lance in Adria, and the Two Crowns count their silver. Me? I just want to get home before the snows bury the graves. And if you hear a cantor singing the \textit{Planctus Armatus}, stop and listen. That's the only honest history left---especially the verse about the queen who married twice, bore sons to both kings, and left the realm wondering which throne was hers by right, and which by rumor.''
\end{quote}

\noindent\textbf{Genre / Mood:} Courtly intrigue, chevauchée warfare, fractured legitimacy, the tragedy of chivalry---a kingdom that cannot mourn because mourning would mean admitting the old world is dead. Beneath the gilt and the heraldry, chivalry is a thin veneer; the iron of siege and sack shows through at every edge.

\noindent\textbf{Starting Location:} The Pont-du-Tithe, its golden arches catching fire in the sunset as pilgrims chant from both banks. A Duchy Constable in splendid armour guards the toll gate. ``The chevauchée burns to the east,'' he says, gesturing toward the Belworth. ``Two Crowns march to meet it. Whose seal do you carry? Speak quickly---the levy waits for no one. And keep your hand away from your sword. The bridge remembers the last blood spilled here. It remembers the Fenwood horses crossing from the east. It remembers the screams.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
Vhasia is a realm of broken coronations and split loyalties, still bleeding from a wound a decade old. Once a proud province of the Utaran empire, the unified Sun Court shattered like its heraldic sunburst when the Fenwoods of Viterra and their allies crossed the Belworth, sacked the capital of Lence, and cast down the old Blood line. Duke Braedyn Fenwood---uncle of Veralyn, now the Warrior Queen of Viterra---led the assault, his ironshod cavalry grinding the Sun King's army into the mud of the Ford of Ashes. They burned Lence until the mirrors melted and the very stones wept. They left the throne empty---not destroyed, but open. And they left a wound that has not healed.

But the wound runs deeper than smoke and stone. Forty years ago, during the Province Wars, King Wymund of Viterra sought alliance with Vhasia through marriage---taking Elianor Everblood (née Longblood) as his queen. She bore him two daughters, his heir Valdais Longblood, and a second son, Wyllmot, who was made heir to the Viterran throne. When Wymund died, Elianor married his son Baldvyn II, becoming queen of Viterra again and binding the two realms in blood---but also in rumor. The jealousies between these unions ignited the provincial wars that weakened both realms, culminating in Wyllmot's clash with Valdais at Adria---a victory stained by the suspicion that brother had fought brother for a throne built on doubtful bloodlines.

When the Fenwoods came, Elianor was in her western duchy of Ethor, having married into the Vhasian royal Bloods line---a final attempt to stitch the realm back together. Her nephew had been made heir to the Sun Court on condition of that marriage, a settlement that the sack of Lence tore apart utterly. Now, Elianor rules from Ethor, her alliances woven from the threads of three marriages and the quiet terror that none of her sons may be truly ``hers'' by the blood that matters. She is not mourning. She is waiting---for proof, for a sign, for the moment when the fractured sun might finally cast a single shadow.

The memory of the sack haunts every negotiation. The Queen-Mother, Elianor Everblood, lost her third husband---the old Sun King---and her cousins and sisters in the flames. Her second son, Wyllmot, King of Viterra and Duke of Aberiel, died fighting rebels in Viterra, leaving his crown to his daughter, Veralyn---who is now the Warrior Queen, who was the Fenwoods' ward, and who was \textit{there} when they burned Lence. Elianor rules from a riverside keep, weaving alliances and hiding a secret---a Dauphin, perhaps her grand-nephew, perhaps a changeling, who may be the key to restoring the Sun Court. Her eldest son, Valdais Longblood, Duke of Adria in the west, claims the throne by blood and steel, but Elianor does not trust him. She has outlived three husbands and two sons. She is not mourning. She is waiting.

The Fenwood sack was not just a military defeat; it was a betrayal of chivalry itself. The knights of Vhasia had sworn to protect the realm with the \textbf{Round of Five Virtues}: Courage, Courtesy, Mercy, Prowess, Truth. At Lence, the Fenwoods showed no mercy. The knights still argue about whether chivalry died there or whether it was already a corpse, dressed in armour, a thin veneer over the cruelty of the world. The Cantor Alain sings the \textit{Planctus Armatus} each spring, and the congregation weeps for Sir Corin, who unbuckled his gorget to hear a thief's voice and was undone by the mercy he offered. Sir Corin died at Lence. He was trying to save a child. The child is now the Dauphin---or so the Queen-Mother claims---and perhaps the living proof that Elianor's bloodline, however tangled, still holds the true sun.

\vspace{2mm}
\textbf{What you notice first:} The smell of old smoke and fresh ink. The way every noble's seal has a hairline crack---like the fault lines in old marriage alliances. The constant, nervous laughter of courtiers who remember whose head rolled last season---and whose mother married which king. The way the knights touch their gorgets when someone speaks of mercy, remembering Sir Corin's unbuckling at Lence---and the queen who watched it all from Ethor, wondering if the mercy she'd been shown was ever real.

\textbf{The one rule that kills newcomers:} Never carry a coin that bears only one sunburst. The realm has three mints, and the wrong metal on the wrong road is treason. Also, never speak the name ``Fenwood'' without naming Queen Veralyn first. The wound is still raw. The Queen-Mother's spies are everywhere---and they have long memories. And never forget: the first treason was not in the field, but in the nursery, when a queen wondered if the son in her cradle was truly the king's.
\end{tcolorbox}

\subsection*{Regional Mood: Crowns, Courts, and the Weight of Lence}

Here, the past is a weapon and the future is a weight---not of silver, but of memory. The Fenwoods came from Viterra across the Belworth with fire and iron, broke the Sun King's army at the Ford of Ashes, and burned Lence until the mirrors melted. Duke Braedyn Fenwood led them; his daughter Veralyn---now the Warrior Queen of Viterra---was her father's attendant and witness to the sack, a fact that complicates every border negotiation. The Vhasian nobility has not forgotten. They cannot forgive. But they cannot afford a war with Viterra, either.

So they fight each other instead.

Now every duke with a long purse and a short conscience backs a claimant. The Queen-Mother weaves alliances from her riverside keep in the west, holding a Dauphin---perhaps her grand-nephew, perhaps a commoner with a convincing birthmark---as a bargaining chip. The Two Crowns---Valdais Longblood of Adria in the west, and a rival claimant backed by the southern parlements---march their levies. Routiers sell their swords by the season, and the memory of Lence is ground into every negotiation. Vhasia is not a nation at war. It is a nation that forgot how to be at peace. And ten years later, the smoke from Lence still hasn't cleared---and neither has the doubt about whose blood truly flows in the veins of those who claim the throne.

\subsection*{The Ninth Taboo Manifestation}
\begin{tcolorbox}[colback=black!3,colframe=red!60!black,title=The Ninth Taboo,breakable]
In any Vhasian court, the ninth citation of precedent voids all previous. The wise stop at eight and settle. Those who cite the ninth are said to hear a faint laughter from the marble walls---and their cases are never heard again. The laughter is the dead king's. He was the ninth king of his line. He died at Lence. He is still laughing---because the ninth citation is the only one that names the Fenwoods as the true heirs. No one has ever cited it. No one dares.

Also, the ninth cup poured at a feast is for the Hollow. The wise pour eight and leave the ninth empty. Those who pour the ninth see a pale figure standing at the high table---waiting to collect a weight. The figure is the dead king, checking his oath-stick.
\end{tcolorbox}

\subsection*{The Province Wars (Forty Years Ago)}

Before the Fenwood sack, there was the \textbf{Province Wars}---a generation of low-intensity raiding between the cantons that left the old crown weakened and the knights bitter. The peace that ended the Province Wars was a treaty of exhaustion, not reconciliation. It was during those wars that the dynastic tangles began: King Wymund of Viterra married Elianor Everblood to seal the peace, only for her to later wed his son Baldvyn II after Wymund's death---a union that sparked the very conflicts it was meant to prevent. The veterans of the Province Wars are now the old men who watch the new claimants with weary eyes. They remember when mercy meant something. They also remember when it cost everything---and they whisper that the first casualty of those wars was trust in the marriage beds of queens.

\subsection*{The Players of the Great Game}

The succession crisis has four major factions, each with a claim, a secret, and a price---each shadowed by the legacy of dual marriages and uncertain parentage.

\begin{itemize}
  \item \textbf{Queen-Mother Elianor Everblood (née Longblood):} Widow of three kings---Wymund of Viterra, his son Baldvyn II, and the last Sun King. Mother of Valdais and the late Wyllmot. She rules the western canton of Ethor. Her price is a guarantee that her grand-nephew (the Dauphin) will not be killed. Her secret: she is not sure if Valdais is truly her son by Wymund---or if he was conceived during the turbulent shift between husbands. She acts as if he is of the Blood, because certainty would destroy what remains of the Longblood line---and because the Dauphin may be her only hope for an heir without doubt.
  \item \textbf{Duke Valdais Longblood of Adria:} The western claimant. Nephew---or half-brother, depending on the rumor---to the last Sun King. A tournament champion with a flawless record and a suspicious habit of opponents falling ill before jousts. His claim rests on blood and steel. His secret: his mother, Elianor, does not trust him---and he suspects why. He was not at Lence. He wishes he had been---so he could have died with honor instead of living with the question: ``Whose son am I, truly?''
  \item \textbf{The Two Crowns (Rival Dukes of the South and East):} Backed by the southern parlements and the eastern march lords who face the Belworth. Their claim rests on election and expediency. Their secret: they have been known to accept Viterran gold---but never the Fenwood seal. They whisper that Veralyn is not her father's daughter. They are wrong. They are also afraid---afraid that if the Viterran line is false, then the Vhasian claim via Elianor's marriages is equally doubtful, and the realm has no true sun left to claim.
  \item \textbf{The Last Dauphin (The False/Real Heir):} A child of nine years, hidden in the Queen-Mother's keep. His legitimacy is a matter of faith. His existence is the only thing keeping Valdais from marching on his mother's castle. His secret: even Elianor does not know if he is truly her grand-nephew---but she has dreamed of Sir Corin, who died saving a child at Lence. The child in the dream has the Dauphin's face---and perhaps, just perhaps, the unbroken line of Elianor's blood, however it flowed.
\end{itemize}

\subsection*{The Cantor Alain and the Planctus Armatus}

No figure embodies Vhasia's tragedy more than \textbf{Alain, the Iron Bard}. A cantor who survived the sack of Lence as a boy---rescued by Sir Corin himself---Alain carries a cracked ring and a voice that carries the weight of the dead. Each dusk, he sings the \textit{Planctus Armatus} (the Lament for the Armoured), a song that recounts Corin's unbuckling and the mercy that killed him. The song has eight verses. The ninth is silence---the space where the Hollow listens. The fourth verse, oft-whispered, speaks of ``the queen who wore two crowns, and the son who knew not which father's blood he carried.''

\textbf{Mechanical Hook:} Once per session, a character who hears Alain sing (or who recites a verse of the \textit{Planctus Armatus}) may gain +1 die on a roll involving mercy or witness. On a failure, they mark 1 Fatigue---the weight of remembering. If they recite the fourth verse specifically, they may instead gain +1 Effect on a roll involving inheritance or legitimacy---but mark 1 Corruption, as the doubt it stirs is a poison to certainty.

\subsection*{Prominent NPCs of Vhasia}
\label{sec:vhasia-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Elianor Everblood & Queen-Mother of Ethor & Restore her grand-nephew (real or false) to the throne---and prove, at last, whose blood truly flows in Valdais's veins. She has outlived three husbands and two sons. She is not mourning---she is planning, and waiting for the moment the doubt ends. Her shadow has begun to drift independently---sometimes toward Ethor, sometimes toward the memory of Wymund's court, sometimes toward the empty space where Baldvyn II's queen should stand. \\
\hline
Valdais Longblood & Duke of Adria, western claimant & Unite Vhasia under a strong western crown---and silence the whispers about his parentage once and for all. He has never lost a joust. His opponents have a habit of falling ill before the tilt. He does not poison them. His mother's spies do. He carries a shard of Wymund's coronation sword in his boot---a reminder that the first throne he ever knew may not have been his by right. \\
\hline
The Red Marshal & Exiled commander of the Sun Court's army & Win back Lence---or die trying & He lost the Battle of Ashes; he still sees it in his dreams and his nightmares. He was the last man to see the old king alive. The king's last words were a curse on the Fenwood name. He wonders, in the quiet hours, if the curse was meant for the Fenwoods---or for the queens who married twice and left the realm questioning whose sons were truly theirs. \\
\hline
Seint Michel's Justiciar & Magistrate of the Order of the Gilded Scale & Enforce the chivalric code without mercy & He carries a scale with one pan empty; the empty pan is for lies. He has never been able to balance it. The empty pan holds the ghost of Sir Corin---and sometimes, he swears, the whisper of Elianor's doubt as she watched the sack from Ethor. \\
\hline
Sister Oria & Custodian of the Lantern-Rites House on the Belworth & Keep the rites of hospitality and witness & She has never drawn a blade. She has never needed to. The Queen-Mother's enemies who cross her threshold have a habit of forgetting why they came. Her scroll records every name and every death of those who fell at Lence. \\
\hline
The Cantor Alain & Iron Bard, bearer of the Cracked Ring & Sing the truth of Corin's mercy before the armour closes again---and sing the truth of the queens who bore sons to two kings, and the sons who never knew which father's blood they carried. Feathers grow from his temples; his shadow moves a half-step behind him. His songs are the only honest history left in Vhasia. He sings the names of the dead every dusk. The dead listen---and sometimes, they answer. \\
\hline
The Dauphin (Child) & A child of nine, hidden in the Queen-Mother's keep & Survive---and perhaps, one day, answer the question that has haunted Elianor since Valdais was in swaddling clothes. He has never seen the sun. He knows three languages, the lineage of every noble house, and the taste of every poison. He does not know if he is a prince or a puppet. He dreams every night of a knight with an unbuckled gorget---and of a queen who wore two crowns, wondering which throne was truly hers. \\
\hline
\end{longtable}

\subsection*{Names of Vhasia}
\label{sec:vhasia-names}

Inspired by the fractured duchies: Francien, Norman, and Brittonic influences. Soft vowels, silent consonants, and surnames that recall a bridge, a ford, or a stained banner. Many names echo the Province Wars era, when marriage alliances were made and broken.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Valdais & Elianor & de la Marche & \textit{the Iron} \\
\hline
Wymund & Baldvyn & l'Ever & \textit{the Unwed} \\
\hline
Renaud & Isabeau & de Lence & \textit{the Unbuckled} \\
\hline
Corin & Oria & d'Ethor & \textit{Bell-Sworn} \\
\hline
Touénot & Mael & du Pont & \textit{the Silver} \\
\hline
Alain & Giselle & de la Croix & \textit{Ash-Banner} \\
\hline
Ghiro & Celandine & l'Argent & \textit{the Unpaid} \\
\hline
\end{longtable}

\paragraph*{(Château/Cathedral/Forest)} Wayside shrine with pilgrims tacking petitions to a scorched sun disc; vine-terrace clos above a millstream with napping watchmen. The oldest vintners whisper that the ``Crownbreaker'' vintage was blessed by Elianor herself---once as Viterran queen, once as Vhasian queen---before the province wars turned the gift to ash.

\section*{Spades --- Places}
\label{sec:vhasia-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Wayside Shrine} -- Pilgrims tack petitions to a scorched sun disc; each parchment curls in different winds.
  \hfill \textit{The petitions are never answered. The shrine is a mailbox for the dead---especially those who fell at Lence. The dead do not write back. They wait.}
\item \textbf{Vine-Terrace} -- Clos above a millstream; watchmen nap, halberds traded for bottles.
  \hfill \textit{The best vintage is called ``Crownbreaker.'' It tastes like smoke and regret. The year of the sack was a particularly good harvest. The wine remembers.}
\item \textbf{Bastide Square} -- Neat as a checkerboard; merchants pay homage to different ``hours'' of the Sun.
  \hfill \textit{The square has a gallows. It rotates to face whichever claimant lost the last duel. The wood is still green---from the forest where the Fenwoods camped before Lence.}
\item \textbf{Royal Forest} -- Antler posts mark laws no king remembers; poachers still hang.
  \hfill \textit{The forest's deer have grown bold. They wear crowns of shed antlers. Some say they are the ghosts of the old nobility---or the knights who unbuckled their gorgets too late.}
\item \textbf{Pont-du-Tithe} -- Tollbridge flanked by chapels and crossbows; chains blessed by rival bishops.
  \hfill \textit{The chains were forged from melted sunbursts. They ring in different keys. The Fenwood army crossed here from the east. The bridge still groans when Viterrans approach. The groans are prayers.}
\item \textbf{Great Fair} -- Outside a chartered town; counting sheds ring like bells, rumors louder still.
  \hfill \textit{The fair has a Court of Miracles where counterfeiters sell to informers. The best forgery is a sunburst coin that rings true until you test its weight. The weight of a lie is memory.}
\item \textbf{Salt Pans} -- Windpumps on an estuary; gulls own the law, smugglers own the night.
  \hfill \textit{The salt is gray. It tastes of old tears and older weights. The smugglers carry more than salt---they carry the Queen-Mother's letters. The letters smell of Lence.}
\item \textbf{Siege-Château} -- Hoardings patched, cistern low, banners cut into ragged suns.
  \hfill \textit{The banner's ragged edges are from a Fenwood sword. The owner still limps. He lost his son at Lence. He fights for Valdais but he does not love him. He loves the ghost of his son.}
\item \textbf{Cathedral Yard} -- Cranes creak, stones fly, scaffolds bristle with rival masons' marks.
  \hfill \textit{The mason's marks are clan signs. A scratched-out mark means a dead foreman. One mason's mark is a sunburst with a crack---the mark of the Queen-Mother's architect. He was at Lence. He built the pyres.}
\item[J] \textbf{Parlement Hall} -- Hung with faded tapestries; sunbursts spliced and restitched by claimants.
  \hfill \textit{The tapestries show battles that never happened---and some that shouldn't have. One tapestry shows the sack of Lence from the Fenwood perspective. The Queen-Mother has ordered it burned three times. It keeps reappearing. The thread is from the shrouds of the dead.}
\item[Q] \textbf{Queen's Causeway} -- Mile-long oath paved in quarried stone; half-sunk each spring.
  \hfill \textit{The sunken half is a graveyard for vows. The stones are slippery with old lies. The Queen-Mother walks here at dawn. She speaks to the dead. The dead answer in the voice of the Dauphin.}
\item[K] \textbf{Sun Palace (Lence)} -- Shuttered mirrors, brambled gardens; courtiers playing at kingship.
  \hfill \textit{The mirrors were covered after the sack. No one wants to see what they show now---a reflection of a kingdom that no longer exists. The courtiers pretend the king still lives. The servants know better. The mirrors know best.}
\item[A] \textbf{King's High Road} -- Milestones chipped of crowns; yet every army still marches it.
  \hfill \textit{The road is paved with the bones of tax collectors. They are not missed. The Fenwood army marched here from the Belworth. Their footprints are still visible---in the stone and in the memory of every orphan.}
\end{enumerate}

\paragraph*{(Lord/Court/Company)} Road warden with dented badge and honest horn; vintner-guild syndic counting saints' days as grace periods.

\section*{Hearts --- People \& Factions}
\label{sec:vhasia-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Road Warden} -- Dented badge, honest horn; ignored by both crowns but knows the roads.
  \hfill \textit{His horn is cracked. It still sounds---but the note is always half a step flat. He was at Lence. He does not speak of it. The horn does.}
\item \textbf{Vintner Syndic} -- Counts saints' days as grace periods, vintages as treaties; wine as diplomacy.
  \hfill \textit{He keeps a bottle of Crownbreaker in every cellar. He knows which vintage each duke prefers. He also knows which duke's wine is poisoned. He does not ask. The poison is memory.}
\item \textbf{Abbess-Chatelaine} -- Rules a fortress-convent with iron rule and sharper scripture.
  \hfill \textit{She has never drawn a blade. The scaffold has, on her orders. She sheltered refugees after the sack. The Queen-Mother owes her a weight. The Queen-Mother does not forget weights---or the smell of burning wool.}
\item \textbf{Routier Captain} -- Scarred and mercenary; prefers ransoms to loyalty; coin as conscience.
  \hfill \textit{He switches banners as often as his shirt. His loyalty is to the paychest---and to his mother. He fought for the Fenwoods at Lence. He still has nightmares. He pays for them in silver.}
\item \textbf{Duchy Constable} -- Splendid armour, empty coffers, pride swollen as a banner; glory over gold.
  \hfill \textit{His armour was his grandfather's. It still has the dent from a Fenwood axe. He serves Valdais because Valdais pays---but he served the old king first. He cannot forgive himself for surviving Lence. The dent weeps rust.}
\item \textbf{Cantor Alain} -- The Iron Bard, bearer of the Cracked Ring; his songs are the only honest history.
  \hfill \textit{Feathers grow from his temples. His shadow moves a half-step behind him. He sings the names of the dead every dusk. The dead listen. And sometimes, they sing back.}
\item \textbf{Salt-Farmer} -- Gabelle lessee with cudgels, writs, and a river-chart no crown has; tax as tyranny.
  \hfill \textit{His tally is written in salt. One drop of water and the obligations vanish---along with the evidence. He is the Queen-Mother's man. He does not say so aloud. His salt tastes of ash.}
\item \textbf{Trouvère} -- A scandalous lay about the last Sun King brings knives to tavern doors.
  \hfill \textit{He sings the truth. That is why he sleeps in a different inn every night. He knows who really ordered the burning of Lence. He will not say. He values his tongue. The truth is a currency he spends in silence.}
\item \textbf{Marshal-Exile} -- The Sun Court's hammer; now wielded by whoever pays in silver.
  \hfill \textit{He lost the Battle of Ashes. He still sees it in his dreams and his nightmares. He was the last man to see the old king alive. The king's last words were a curse on the Fenwood name. The curse is still unpaid.}
\item[J] \textbf{Heretic Perfect} -- Preaches peace while tucking hidden daggers into parchment margins.
  \hfill \textit{His sermon is a code. The psalms tell you which bridge to cross and which to burn. He is the Queen-Mother's spy. He is also Valdais's spy. He is also his own spy. His sermons are the only honest prayers left.}
\item[Q] \textbf{Queen-Mother} -- Riverside keep in the west; weaves alliances as if they were dowries.
  \hfill \textit{She has outlived three husbands and two sons. She is not mourning. She is waiting. Her grand-nephew---real or false---will sit on the Sun Court. She will not live to see it. She has made her peace with that. The peace tastes of Lence.}
\item[K] \textbf{Two Crowns} -- Rival dukes, each ``crowned'' in rival cathedrals, neither whole.
  \hfill \textit{One wears a crown of iron. The other wears a crown of words. Both rust. Neither was at Lence. Neither speaks of it. They pay others to speak for them. And sometimes, they whisper that if the Viterran line is false, then the Vhasian claim via Elianor's marriages is equally doubtful.}
\item[A] \textbf{Last Dauphin} -- A rumor with a birthmark; armies march for whispers alone.
  \hfill \textit{He may be a baker's son. The armies don't care. They need a sun to follow. He has never seen the sun. He knows three languages, the lineage of every noble house, and the taste of every poison. He does not know if he is a prince or a puppet. He does not ask. He dreams every night of a knight with an unbuckled gorget.}
\end{enumerate}

\paragraph*{(Chevauchée/Law/Church)} Chevauchée burning loyalty oaths as well as barns; smoke curls like scripture. Interdict falls on a county; bells silenced, trade stalls.

\section*{Clubs --- Complications/Threats}
\label{sec:vhasia-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Chevauchée} -- Raiders burn loyalty oaths as well as barns; smoke curls like scripture.
  \hfill \textit{The raiders wear no colors. Their horses answer to whistles and grudges. They were at Lence. They remember. The smoke smells of old wool.} `[WAR]`
\item \textbf{Interdict} -- Falls on a county; bells silenced, trade stalls, heretics prosper.
  \hfill \textit{The bells were melted for shot three wars ago. The silence is old. The interdict was ordered by the Cardinal-Regent---but no one knows which claimant he serves. The silence is a prayer for the dead of Lence.} `[FAITH]`
\item \textbf{Forest Law} -- Venison becomes treason; poachers hang from antler posts.
  \hfill \textit{The antler posts are carved with the names of the hanged. They whisper at dusk. The names include survivors of Lence who could not find work.} `[DEATH]`
\item \textbf{River Spate} -- Ferries chained, bridges drowned, pilgrims stranded mid-prayer.
  \hfill \textit{The water is the color of old gold. It washed away a mint last spring. The mint belonged to a claimant who could not pay his routiers. The routiers now raid for the other side.} `[WEATHER]`
\item \textbf{Coin Debasement} -- Three mints, three suns, no trust; soldiers demand silver alone.
  \hfill \textit{The silver is from the same mine---but the stamp is a prayer, not a promise. The Queen-Mother controls two mints. Valdais controls one. The Cardinal-Regent holds the dies. The dies were made from melted Fenwood spurs.} `[WEIGHT]`
\item \textbf{Parlement Divided} -- A siege laid on parchment before walls; war by paperwork.
  \hfill \textit{The clerks duel with citations. The loser pays the winner's ink bill. The winner gets to draft the next round of tax levies. The peasants pay both. The ink is mixed with ash from Lence.} `[LAW]`
\item \textbf{Company Mutiny} -- Unless arrears are paid in blood or coin; mercenaries as a time bomb.
  \hfill \textit{The mutineers have a list. Your name is on it---or will be by nightfall. The list was leaked by the Queen-Mother's spy. She wants the routiers to switch sides. The list is written on a page torn from the Chivalric Roll.} `[MERCENARY]`
\item \textbf{Relic Dispute} -- Two abbeys claim the same saint's bones; mobs answer with stones.
  \hfill \textit{The bones are actually from a cow. The saints don't mind. They like the attention. The abbeys are backed by rival claimants. The cow was from a farm that burned at Lence.} `[FAITH]`
\item \textbf{Harvest Blight} -- Bread locked behind pikes; bakers guarded as lords.
  \hfill \textit{The blight has a name. It is the name of a duke who raised the levy too late. His fields fed the Fenwood army. His people starved. He still rules. The blight is his guilt made visible.} `[DISASTER]`
\item \textbf{Tournament ``Accident''} -- Hides an assassination; your favor is now evidence.
  \hfill \textit{The lance was filed with a prayer. The prayer was answered---but not for the target. The victim was a claimant's champion. The killer was never found. The Queen-Mother smiled. The lance was from Lence.} `[COMBAT]`
\item[J] \textbf{Royalist Restoration} -- Sunburst badges painted overnight on doors and graves.
  \hfill \textit{The paint is the color of dried blood. The badges face the wrong direction. The Queen-Mother's agents paint them. She wants the people to remember that the Sun Court once existed. The badges glow in moonlight.} `[OMEN]`
\item[Q] \textbf{Feudal Levy} -- Lords demand levies in days, not months, with three days' bread.
  \hfill \textit{The bread is hardtack from the last Fenwood siege. It still smells of smoke. The levy was called by Valdais. The lords are his cousins. They do not want to march. They owe him.} `[WAR]`
\item[K] \textbf{Winter Campaign} -- Roads turn to mud, hooves to ice, armies to ruin.
  \hfill \textit{The snow is gray with ash. The Pass of Ashes has a sister pass now. The Fenwood army marched through the Pass of Ashes from the Belworth. The ghosts still march there---in both passes.} `[WEATHER]`
\item[A] \textbf{Lence's Echo} -- Smoke rises from the ruins. The Fenwood banners appear on the horizon. It is a false alarm---or is it?
  \hfill \textit{The party must test Resolve (DV 4) or lose 1d4 Fatigue from remembered terror. If they fail, they also gain the \textbf{Haunted} condition---the dead of Lence will follow them for one session. The question they whisper is always the same: ``Why did you survive?''} `[FEAR]` `[DEATH]`
\end{enumerate}

\paragraph*{(Charter/Patent/Right)} Safe-conduct sealed by a duke or abbey; honoured only once, on one road; burgess charter for a market ward.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:vhasia-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Safe-Conduct} -- The seal is cracked. The road is watched. Use it before the seal breaks.
  \hfill \textit{The seal was cracked at Lence---by a Fenwood axe.} `[TRAVEL]`
\item \textbf{Burgess Charter} -- The charter is a bell. Ring it, and the town answers---or it doesn't.
  \hfill \textit{The bell was cast from a melted sunburst. It rings true---but only for those who remember. The bell's clapper is a finger bone from a knight who died at Lence.} `[SOCIAL]`
\item \textbf{Bridge Farm} -- Tolls at Pont-du-Tithe for a season, if you can hold it.
  \hfill \textit{The toll-keepers are loyal to the last coin. Yours---or theirs. The bridge was rebuilt after Lence. The stone is new. The ghosts are old. The ghosts do not pay tolls.} `[TRADE]`
\item \textbf{Pareage Charter} -- Shared rule with a monastery; half justice, half rent, no peace.
  \hfill \textit{The monks pray for you. They also pray for your enemies. Balance. The monastery sheltered refugees after the sack. The Queen-Mother has not forgotten.} `[LAW]`
\item \textbf{Wardship} -- Over a minor heir; lands obey you ``for now.''
  \hfill \textit{The heir has a knife. He knows how to use it. He learned from watching you. His father died at Lence. He does not know which side his father fought for. He does not care. The knife was his father's.} `[AUTHORITY]`
\item \textbf{Gabelle Lease} -- Salt-tax rights; lucrative, hated, and dangerous.
  \hfill \textit{The lease is a death warrant written in profit. Cash it wisely. The salt pans are near the Belworth. The Viterran border is close. The Fenwood raids come at dawn from the east. The salt is gray with ash.} `[TRADE]`
\item \textbf{Letters Patent} -- Raise a free company under your banner; an army as a business.
  \hfill \textit{The letters are signed in blood---yours or the notary's. Both are valid. The notary lost a hand at Lence. He signs with his left. His right is buried in Lence.} `[MERCENARY]`
\item \textbf{Remission} -- Of feudal dues for a village; they'll march when called.
  \hfill \textit{The village has a long memory. They remember the last lord who promised mercy. He was a Fenwood. They do not forgive. The remission is written on a page torn from the Chivalric Roll.} `[COMMUNITY]`
\item \textbf{Low-Justice Grant} -- Stocks, fines, gallows in a troublesome hamlet.
  \hfill \textit{The gallows rope is new. The notches are from the last hamlet. The hamlet was burned by routiers. The routiers were paid by a claimant. The claimant denies it.} `[LAW]`
\item[J] \textbf{Confiscation Writ} -- Against a royalist estate---if you can seize it first.
  \hfill \textit{The writ is a race. The other claimant rides a faster horse. The estate belonged to a lord who died at Lence. His widow still lives there. She has a knife. The knife is for the horse.} `[WEIGHT]`
\item[Q] \textbf{Queen's Audience} -- Private with the Queen-Mother; a whisper becomes policy.
  \hfill \textit{The audience is a chess game. She always wins---but she lets you take a pawn. The pawn is the Dauphin. Do not touch him. His shadow is not his own.} `[AUTHORITY]`
\item[K] \textbf{Sunburst Warrant} -- Of the fallen court; most still bow if shown boldly.
  \hfill \textit{The warrant is a ghost. It has no power---only the power of belief. The ghost was signed by a dead king. The dead king's seal was stolen from his tomb. The Queen-Mother has the original. The original weeps wax.} `[LEGEND]`
\item[A] \textbf{General Pardon (Temporary)} -- Rebels absolved if they bend the knee.
  \hfill \textit{The pardon is a spring. It will close when the season turns. The rebels were at Lence. They fought for the old king. They fled when the Fenwoods broke the line. They are still running. The pardon smells of ash.} `[JUSTICE]`
\end{enumerate}

\section*{Quick use notes}
\label{sec:vhasia-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer (2--5 → 4, 6--10 → 6, J/Q/K → 8, A → 10).
\item Diamonds are codified outcomes (charters/patents/rights) that change position rather than call for a roll.
\item If any A appears, echo \textbf{fractured sun motifs}---light that splinters, oaths that crack like glass, smoke that smells of Lence, and a child who may be a prince or a puppet. Also echo the weight of memory: a ghost knight, a mirror that shows the sack, a bell that rings for the dead.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:vhasia-mechanics}

\begin{itemize}
\item \textbf{Fractured Seals:} When brandishing a writ, badge, or charter tied to a Sun Court claimant, step Position up one step for the scene's next interaction at a gate, court, or checkpoint. Each use advances a \textbf{Legitimacy [4]} timer linked to the claimant; when it fills, rivals challenge the authority on sight. The seal's crack widens with each challenge. Eventually, it will break---and the broken seal will smell of Lence.
\item \textbf{Parlement Games:} If a scene touches petitions, guild disputes, or feudal appeals, the first Diamond played becomes a policy lever: treat it as two Diamonds toward swaying a decision or rewriting a custom. Afterward, start or advance a \textbf{Whispered Precedent} timer; when it fills, backlash spreads as other courts cite the case against you. The Queen-Mother's clerks are masters of this. They have never lost---because they have never forgotten the price of losing. The price is Lence.
\item \textbf{Gabelle Knife-Edge:} At bridges, fairs, or salt depots, convert Diamonds tied to tolls, leases, or safe-conducts into immediate passage or tariff relief for the company (no roll). Each conversion ticks a \textbf{Resentment [4]} timer for the locale; when it completes, expect boycotts, ambush rumors, or an audit by rival officials. The salt farmers remember Lence. They do not forget. The salt in their tears is gray.
\item \textbf{Vhasian Chivalry \& the Round of Five Virtues:} The knights of the western cantons (Ethor, Adria) follow the \textbf{Round of Five Virtues}: Courage, Courtesy, Mercy, Prowess, Truth. In a formal joust or a legal duel, the GM may award \textbf{Victory by Virtue} rather than by blood: the combatant with the higher Virtue score wins the dispute, even if they lost the physical exchange. This is the law of the tilt and the law of the court. But at Lence, the Fenwoods showed no mercy. The knights still argue about whether chivalry died there---or whether it was already a corpse, dressed in armour, a thin veneer over savagery. And sometimes, they whisper that the first crack in the sun came not from the Fenwoods' swords, but from the marriage beds where queens bore sons to two kings---and the sons who never knew which father's blood they carried.
\item \textbf{The Armor Closes / The Song Escapes (Chivalric Timers):} The Cantor Alain, bearer of the Cracked Ring, sings a warning at every tourney:
  \begin{itemize}
    \item \textbf{Armor Closes [6]} advances when knights choose protection over mercy, when gorgets are buckled tight, when chivalry becomes siege mentality. When full, the cantons close their borders and the raids become wars. The Fenwoods taught Vhasia that mercy is a weakness. The knights have not forgotten. They wear their gorgets to bed.
    \item \textbf{Song Escapes [8]} advances when a Cantor sings the truth, when mercy is shown, when armour is unbuckled. When full, a moment of genuine connection occurs---though it may be fatal. The Cantor Alain has been trying to fill this timer for ten years. He is still trying. He sings every dusk. The dead listen---and sometimes they answer.
  \end{itemize}
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Seint Michel, the Judge of the Scar (Mykkiel):} The law that does not bend; the judge who cannot be bought. His shrines are at tourney fields and courtrooms. Knights swear oaths on his scale before the joust.
    \item \textbf{Seint Isabeau, Who Slumbers Amongst the Lilies (Ikasha):} The dream of setting aside the crown; the mercy of forgetting duty. Her groves demand that armour be unbuckled at the threshold.
    \item \textbf{Seint Evroul of the Cairn (Mor'iraath):} The oath that outlasts the oath-taker; the blood-price paid in stone. The barrow-kings demand tribute in stories, not gold.
    \item \textbf{The Cantor's Road (The Traveler):} The road that belongs to no lord; the path that finds its own way. Cantors are immune to chivalric protection---but subject to its violence.
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross-Expansion Hooks)}
\label{sec:vhasia-lore}

\begin{itemize}
  \item \textbf{Seint Michel, the Judge of the Scar (Mykkiel):} The Order of the Gilded Scale maintains the \textbf{Chivalric Roll}---a record of every oath sworn and broken in the cantons. A page torn from this roll can serve as a single-use Claim String (gain +1 Position in any court). The Order's justiciars are feared and respected; they never unbuckle their gorgets, and they have never lost a case. The Queen-Mother keeps a copy in her riverside keep, wrapped in a gray shroud. The page that records the Fenwood oath of safe passage---broken at Lence---is still blank. The Order will not write it. They are afraid of what will happen if they do. The blank page smells of smoke.
  \item \textbf{Seint Isabeau, Who Slumbers Amongst the Lilies (Ikasha):} In the Ethorian marches, knights whisper prayers to Seint Isabeau when they long to set aside their steel. Her shrine is a hawthorn grove where armour must be left at the gate. Pilgrims who enter bare-skinned and speak a true regret find that their next wound heals in a single night---but they dream of a woman in white who asks, ``Will you remember how to be touched?'' Those who return to the grove a second time are never seen again. The knights who survived Lence do not go near the grove. They are afraid of what they might forget---or what they might remember.
  \item \textbf{Seint Evroul of the Cairn (Mor'iraath):} In the high passes of the Ethorian marches, a rusted knight in an unopened helm haunts the old barrows. He is the patron of those who carry a blood-weight and cannot rest. To call upon him, one must pour a cup of salt water at a cairn and name the wrong done to your house. He will witness one blow; if it lands, the foe is \textbf{Marked by the Cairn}---any spirit may locate them for a year. If it misses, the blow turns inward. The Marshal-Exile carries a scar from such a bargain; he does not speak of it. He named the Fenwoods. The blow missed. He still does not understand why. The Cairn weeps rust.
  \item \textbf{The Order of the Sunken Cup (The Hospice of the Broken Blade):} A gray-robed hospice in the ruins of Lence treats the wounded of both crowns. They ask no payment, but they record every name in a scroll that never dries. The Restorationists suspect they are spies for the Queen-Mother; the Two Crowns accuse them of erasing obligations. Sister Oria, the hospice's custodian, is also the keeper of the Lantern-Rites House on the Belworth---neutral ground where even claimants must leave their blades at the door. The scroll has a page for every person who died at Lence. The page is still being written. The ink is still wet. The ink is made from ash and tears.
  \item \textbf{The Ninth:} (See the Ninth Taboo Manifestation box above.)
\end{itemize}

\subsection*{Vhasian Superstitions (burn for +1 die once)}
\begin{itemize}
  \item \textbf{Bread to the Sun:} Crumble a crust onto a cracked sun disc. Ignore the first \emph{Chevauchée} penalty. \hfill `[SUPERSTITION]` \textit{The sun disc should be from Lence. The older the crack, the better. The bread should be stale---like the memory of the old king.}
  \item \textbf{Knot the Seal:} Tie a single knot in a charter's ribbon. Cancel \emph{Parlement Divided} or take --1 die on your next legal argument (your choice). \hfill `[SUPERSTITION]` \textit{The ribbon should be black. Black is for mourning. The knot should be tied while facing east---toward Lence and the Belworth.}
  \item \textbf{Name to the Salt:} Whisper a dead claimant's name into a salt pan. Gain +1 Effect when collecting from their faction this scene, but mark Obligation to the Queen-Mother. \hfill `[SUPERSTITION]` \textit{The salt pan should be on the Belworth. The dead claimant's name should be spoken only once. The Queen-Mother will know. The salt will taste of ash.}
  \item \textbf{The Lence Candle:} Light a candle in the ruins of Lence. For the next scene, any oath sworn by candlelight is binding---but the candle will gutter if the oath is false. The dead will witness. They do not forgive. \hfill `[SUPERSTITION]` \textit{The candle should be made from the fat of a knight who died with his gorget buckled. There are many.}
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:vhasia-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Rival officials or court guards take notice.
\item \textbf{Noise:} Parlement clerks or routier patrols are alerted.
\item \textbf{Trace:} Trackers or tax collectors mark your route.
\item \textbf{Delay:} A setback costs time or a favourable legal window.
\item \textbf{Supply Strain:} Mark +1 segment on a relevant resource timer.
\item \textbf{Lence Echo:} A flash of smoke, a distant scream, a child's face in a window. The memory of the sack surfaces. The GM gains 1 SB to spend on a fear effect. The fear is always the same: that it will happen again.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} A duchy constable or abbess-chatelaine begins responding.
\item \textbf{Position Lost:} Legal challenge or road closure costs you advantage.
\item \textbf{Foe Appears:} A rival claimant or free company arrives.
\item \textbf{Gear Trouble:} A piece of equipment becomes Compromised/Neglected.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test.
\item \textbf{Claimant's Agent:} A spy working for Valdais, the Queen-Mother, or the Two Crowns reveals themselves. They demand a favor or a bribe. The bribe must be a memory of Lence. If you have none, they will take a finger.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional routiers, parlement guards, or ducal forces arrive.
\item \textbf{Key Gear Breaks:} A crucial tool or weapon becomes temporarily unusable.
\item \textbf{Major Twist:} A writ is voided, an interdict declared, or a levy called.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
\item \textbf{Condition Applied:} Mark Fatigue 1/Harm 1 or an appropriate Condition.
\item \textbf{The Dauphin's Name:} Someone speaks the child's name aloud. The Queen-Mother's spies are everywhere. The party is now on her radar. The child will dream of them.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates fully.
\item \textbf{Authority Arrival:} The Two Crowns, Queen-Mother, or Marshal-Exile intervenes.
\item \textbf{Scene Shift:} A chevauchée hits, the river floods, or the parlement erupts.
\item \textbf{Patron Omen:} An omen appears, a blessing is lost, or a curse manifests.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\item \textbf{The Sack Relived:} Smoke rises from Lence. The Fenwood banners appear on the horizon. It is a false alarm---or is it? The party must test Resolve (DV 4) or lose 1d4 Fatigue from remembered terror. If they fail, they also gain the \textbf{Haunted} condition---the dead of Lence will follow them for one session.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Fractured Authority:} Seals break mid-signing, legal papers rewrite themselves, badges refuse honor.
\item \textbf{Legal Warfare:} Court decisions reverse, precedents are cited against you, writs expire prematurely.
\item \textbf{Sun Symbolism:} Sun discs crack underfoot, light splits into contradictory shadows, oaths shatter like glass.
\item \textbf{Lence's Shadow:} A ghost appears---a soldier, a child, a king. It asks a question. Answering costs a memory. Refusing costs a curse. The question is always: \textit{``Why did you survive?''}
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Dust Cloud}.
\item \textbf{Cut the Charter (Presence + Sway):} On a hit, freeze range for one exchange; if a Safe-Conduct is in your possession, +1 die.
\item \textbf{Weather the Sack (Resolve + Spirit):} On a hit in \emph{Lence's Echo}, you pick who keeps their memories; on a miss, the GM does.
\end{itemize}

\subsection*{Boss Seeds (three-phase foes)}
\begin{itemize}
  \item \textbf{The Red Marshal (Exiled Sun Court hammer)} -- \emph{Phases}: Chevauchee Call $\rightarrow$ Banner Turn $\rightarrow$ Lists to War. \emph{Cracks}: Letters Patent buys one round of obedience; General Pardon lets a unit stand down. He was at Lence. He ran. He has not forgiven himself. He wears his gorget unbuckled.
  \item \textbf{Cardinal-Regent of the Broken Sun (scarlet gloves)} -- \emph{Phases}: Interdict Blade $\rightarrow$ Seal Both Ways $\rightarrow$ Confessional Leak. \emph{Cracks}: Queen's Audience cancels Interdict Blade once; a Synod Indulgence treats a church check as controlled for one scene. He was not at Lence. He blessed the Fenwood banners before they marched. He does not speak of it. His gloves are scarlet for a reason.
  \item \textbf{The Parlement Spider (chief clerk; quills like daggers)} -- \emph{Phases}: Motion to Delay $\rightarrow$ Counter-Citation $\rightarrow$ Filed at Dawn. \emph{Cracks}: Sunburst Warrant opens a ``royal window''; a Burgess Charter lets a town court overrule once locally. He keeps the lists. He knows which nobles voted for the Fenwood peace. He does not show the list to anyone. The list is written on his own skin.
  \item \textbf{The Two Crowns (rival dukes)} -- \emph{Phases}: Contradictory Edicts $\rightarrow$ Levy in Hours $\rightarrow$ Coin Trial. \emph{Cracks}: Sunburst Warrant $\rightarrow$ public mass unites them for one scene; General Pardon calms levy unrest once. They were not at Lence. They sent levies. The levies never returned. They pay the widows' pensions. They do not visit the graves. The graves visit them.
  \item \textbf{The Queen-Mother (Elianor Everblood)} -- \emph{Phases}: Weave the Web $\rightarrow$ Unveil the Dauphin $\rightarrow$ The Sun Rises Again. \emph{Cracks}: produce proof of the Dauphin's illegitimacy, convince Valdais to kneel, or find the old king's true heir (a child hidden in the Mistlands). She has outlived three husbands and two sons. She will outlive you. Her webs are woven from the hair of the dead.
  \item \textbf{The Cantor Alain (Iron Bard, bearer of the Cracked Ring)} -- \emph{Phases}: The Planctus $\rightarrow$ The Witness $\rightarrow$ The Silence. \emph{Cracks}: unbuckle a gorget in his presence, speak a true regret aloud, or sing the ninth verse---the one that names the Hollow. He is not an enemy. He is a mirror. Do not break the mirror.
\end{itemize}

\subsection*{Legendary Sites \& Trials}
\begin{itemize}
  \item \textbf{Tournament Ordeal (Bastide Square)} -- Declare Purse, Patent, or Pardon as stakes. Win to cash a matching $\diamond$; lose and spawn \emph{Tournament ``Accident''}. The square was built on the site of a mass grave from Lence. The dead watch. They are not impartial. They vote with the wind.
  \item \textbf{Cathedral Scaffold Assize} -- Argue law atop creaking beams; each good point moves Position up; each slip marks Exposure. A Paréage Charter lets you call a recess once. The cathedral was rebuilt after the sack. The old stones are in the foundation. They still remember the fire. They also remember the screams.
  \item \textbf{Pont-du-Tithe Vigil} -- Keep the bridge peaceful through dusk: clear 3 segments on Bridge Peace by feeding pilgrims, quieting relic zeal, and paying one tithe from your purse. The bridge was the first structure the Fenwoods burned. The replacement stone was brought from Viterra across the Belworth. It still smells of smoke---and of Veralyn's perfume. She was there.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power},breakable]
In Vhasia, power flows through the remnants of imperial authority, fractured legitimacy, and the careful manipulation of memory and mercy. Former Utaran provincial structures still provide a framework for governance, but competing claimants to the Sun Court have created a complex web of overlapping jurisdictions and conflicting loyalties. The true power lies with those who can navigate this legal maze---and who understand that the only currency that matters is the memory of Lence.

But beneath the charters and the writs runs a deeper current: the wound that will not close. The Fenwood sack was not just a military defeat; it was a wound to the soul of Vhasia, proof that chivalry was only ever a thin veneer. Every negotiation, every duel, every sealed document carries the weight of that loss. The Queen-Mother plays a long game, her Dauphin a card she may never play. Valdais wants the throne his father could not hold. The Two Crowns will accept any solution that leaves them richer. And the common folk---the farmers, the salt-panners, the road wardens---just want the raiding to stop.

They will wait. They have waited ten years. They can wait longer.

\textbf{For the GM:} Patronage in Vhasia revolves around legitimacy, legal documents, and the ability to command respect from institutions that no longer have a clear central authority. Rewards often take the form of charters, writs, and official positions that can be leveraged into greater influence. To emphasize the post-sack grief and the fragile peace:
\begin{itemize}
  \item Tie rewards to visible symbols (seals, badges, writs) that can be challenged, stolen, or voided---especially if they bear the old king's seal or the Fenwood mark.
  \item Let rival claimants issue conflicting authorities, forcing players to choose whose favor matters more---and whose memory of Lence they are willing to invoke.
  \item Use the parlements, cathedrals, and châteaux as arenas for social contests, where knowledge of law and custom determines success---and where the mere mention of the Fenwood name can shift the tide.
  \item Introduce the \textbf{Lence's Shadow} mechanic: whenever a scene involves a legal dispute, a military manoeuvre, or a negotiation with a noble who lost family in the sack, the GM may spend 1 SB to invoke the memory of the burning. The Position worsens by one step, and the losing party may gain the \textbf{Haunted} condition.
  \item Use the \textbf{Armor Closes} and \textbf{Song Escapes} timers as a visible campaign front. The party's actions---acts of mercy or cruelty---tick one or the other. When the Armor Closes, the cantons become fortresses; when the Song Escapes, a moment of genuine connection may change everything---or kill everyone.
\end{itemize}
In Vhasia, your papers are your power, and your power determines whether you rule or are ruled. But your memory of Lence---and how you use it---determines whether anyone will follow you.
\end{tcolorbox}

\subsection*{The Vacant Throne \& Campaign Timers (For longform political campaigns)}
\label{sec:vhasia-campaign-timers}

\begin{itemize}
  \item \textbf{Vacant Throne Timer [10]:} Advances when a claimant gains a major advantage (battle, noble defection, foreign alliance). Retreats when the party brokers a truce or exposes a weakness. When full, a claimant takes Lence and declares themselves king. The Queen-Mother will not live to see it. She will die the day the sun rises whole again.
  \item \textbf{Armor Closes [6]:} Advances when knights choose protection over mercy, when gorgets are buckled tight, when chivalry becomes siege mentality. When full, the cantons close their borders and the raids become wars. The Fenwoods taught Vhasia that mercy is a weakness. The knights have not forgotten. They sharpen their gorgets.
  \item \textbf{Song Escapes [8]:} Advances when a Cantor sings the truth, when mercy is shown, when armour is unbuckled. When full, a moment of genuine connection occurs---though it may be fatal. The Cantor Alain has been trying to fill this timer for ten years. He is still trying. He sings every dusk. The dead listen---and sometimes they answer.
  \item \textbf{Claimant's Coup (per claimant):} A 6-timer that advances when that claimant wins a Parlement vote, a joust, or a skirmish. When full, they march on Lence. The Dauphin---real or false---watches from the Queen-Mother's keep. He does not know if he should cheer or weep.
  \item \textbf{Lence's Reckoning [8]:} Tracks the party's engagement with the memory of the sack. Advances when they aid a survivor, recover a relic, or speak the truth about what happened. When full, a survivor reveals a secret: the old king's last words, the location of a hidden treasure, or the true parentage of the Dauphin. The secret will change the succession. It will also change the party. Some secrets are heavier than crowns.
\end{itemize}

% Index entries
\index{Vhasia|see{Fractured Sun}}
\index{Theme generator}
\index{Fractured Sun generator}
\index{Places!Vhasia}
\index{People!Vhasia}
\index{Complications!Vhasia}
\index{Rewards!Vhasia}
\index{Queen-Mother Elianor}
\index{Valdais Longblood}
\index{Cantor Alain}
\index{Sir Corin}
\index{Lence}
\index{Fenwood sack}
\index{Province Wars}
\index{Round of Five Virtues}
\index{Armor Closes timer}
\index{Song Escapes timer}
\index{Chivalric Roll}
\index{Seint Michel}
\index{Seint Isabeau}
\index{Seint Evroul}
\index{Local Patrons!Vhasia}
\subsection*{The Succession Mini-Campaign: The Dauphin's Truth}
\label{sec:vhasia-succession}

The central question of any Vhasian campaign is whether the Dauphin is truly of the Blood. This is not predetermined. Each campaign generates its own truth.

\subsubsection*{Determining the Dauphin's Legitimacy (GM Only)}

At the start of the campaign, roll 1d6 on the table below. The result determines the Dauphin's true nature—and the secret the Queen-Mother has been keeping.

\begin{center}
\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[3]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
d6 & Legitimacy & The Truth \\
1 & True Heir & The Dauphin is exactly who the Queen-Mother claims: her grand-nephew, the last of the old Sun King's line. The blood is true. The problem is that Valdais's claim is also true, and the realm must choose. \\
2 & False Heir (Known) & The Queen-Mother knows the Dauphin is a commoner—a Tulkani orphan with a convincing birthmark. She maintains the fiction because without a claimant, the realm falls to Valdais, whom she does not trust. The child does not know. \\
3 & False Heir (Unknown) & The Dauphin is a changeling, placed in the cradle by agents of the Ninth to destabilize the succession. The Queen-Mother suspects but cannot prove it. She has dreamed of Sir Corin, who died saving a child at Lence—and the child in the dream has the Dauphin's face. \\
4 & Uncertain & The Queen-Mother genuinely does not know. The child was brought to her by a dying servant who claimed he was the Dauphin. She has raised him as her own, hoping the doubt will never be tested. She has outlived three husbands and two sons. She cannot bear to lose a fourth. \\
5 & Double Claim & There are two Dauphins. A second child, hidden in the Mistlands, also bears the birthmark and the blood-right. The Queen-Mother knows of this child but has not acknowledged them. The realm may split again. \\
6 & The Blood Remembers & The Dauphin carries the blood of the old Sun King—but also the blood of a Fenwood. The sack of Lence was not just a conquest; it was a family feud. The child is the living proof of a union that should never have happened. \\
\end{tblr}
\end{center}

\subsubsection*{The Succession Campaign Structure}

A Vhasian succession campaign follows three phases. The Dauphin's legitimacy, determined above, shapes every encounter.

\paragraph{Phase 1: The Investigation (Sessions 1--3)}
The party is hired (by the Queen-Mother, by Valdais, or by a neutral party) to determine the truth of the Dauphin's claim. Key scenes:

\begin{itemize}
  \item \textbf{The Queen-Mother's Keep:} Elianor grants an audience. She speaks of the Dauphin but does not show him. She tests the party's loyalty with small favors. Her spies watch.
  \item \textbf{The Cantor's Testimony:} Alain sings the \textit{Planctus Armatus}. If the party listens carefully, the fourth verse hints at the Queen-Mother's doubt. If they press him, he may reveal the cracked ring—and the name of the child Sir Corin died to save.
  \item \textbf{The Valdais Gambit:} Valdais offers the party a counter-offer: find proof that the Dauphin is false, and he will grant lands, titles, and his eternal gratitude. He is desperate. He does not know the truth himself.
  \item \textbf{The Two Crowns' Proposal:} The rival dukes offer a simpler bargain: support their election, and the Dauphin's legitimacy becomes irrelevant. They will pay in silver and pardons.
\end{itemize}

\paragraph{Phase 2: The Confrontation (Sessions 4--6)}
The party has gathered evidence. Now they must decide:

\begin{itemize}
  \item \textbf{The Dauphin's Reveal:} The child is brought before the party. He is nine years old. He knows the taste of every poison. He has never seen the sun. He asks: ``Am I real?''
  \item \textbf{The Trial of Blood:} A formal inquiry is called at the Cathedral Yard. Claimants present evidence. The party may testify. The dead of Lence are watching.
  \item \textbf{The Forged Charter:} Someone has produced a document that purports to prove the Dauphin's legitimacy—or his falsehood. The document is a forgery. Who made it? The party must decide whether to expose it or use it.
  \item \textbf{The Queen-Mother's Confession:} Elianor speaks privately to the party. She reveals the truth (as determined by the d6 roll). She asks them to keep the secret—or to act on it.
\end{itemize}

\paragraph{Phase 3: The Resolution (Sessions 7--8)}
The succession is decided. The party's actions determine the outcome:

\begin{itemize}
  \item \textbf{The Vacant Throne Fills:} One claimant takes Lence. The Queen-Mother may live or die. The Dauphin may rule, be exiled, or be killed.
  \item \textbf{Armor Closes or Song Escapes:} The party's choices have pushed one of the chivalric timers to full. The cantons either seal their borders or open their hearts. The realm changes.
  \item \textbf{Lence's Reckoning:} The memory of the sack is either honored or buried. A survivor reveals a final secret—the old king's last words, the location of a hidden treasure, or the true name of the child who was never meant to be found.
\end{itemize}

\subsubsection*{The Dauphin's Fate (Resolution Table)}
When the succession is decided, roll or choose the Dauphin's fate based on the party's choices:

\begin{center}
\begin{tblr}{
  colspec = {X[2] X[3] X[2]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Outcome & Dauphin's Fate & Campaign Impact \\
Valdais Wins & The Dauphin is exiled to the Mistlands (or killed). & Vhasia unites under a strong western crown—but the Queen-Mother is dead or imprisoned. The memory of Lence fades, but the suspicion of illegitimacy remains. \\
Two Crowns Win & The Dauphin becomes a puppet figurehead. & Vhasia is ruled by a council of dukes. The realm stabilizes but remains divided. The Dauphin grows up in a gilded cage, learning to wait. \\
Queen-Mother Triumphs & The Dauphin is crowned, with Elianor as regent. & Vhasia returns to the old ways. The Queen-Mother dies within a year. The Dauphin's legitimacy is still questioned, but his reign begins. \\
Party Reveals Falsehood & The Dauphin is revealed as a fraud. & The realm descends into chaos. Valdais and the Two Crowns fight openly. The party may be hailed as heroes—or blamed for the war. \\
Party Protects the Secret & The Dauphin rules, but the secret remains hidden. & Vhasia is stable, but the hollow is watching. The party holds a weight they cannot share. The truth may surface at the worst possible moment. \\
The Ninth Intervenes & The Dauphin's legitimacy is erased from all records. The Hollow forgets. & The succession is voided. A new claimant must be found. The Ninth's Coterie gains influence. The party must decide whether to restore the records or let them lie. \\
\end{tblr}
\end{center}

\subsubsection*{GM Guidance: Running the Succession}

\begin{itemize}
  \item \textbf{Let the dice decide:} The Dauphin's legitimacy is random each campaign. This prevents players from assuming they know the outcome. The mystery is real.
  \item \textbf{The Queen-Mother's knowledge:} If the Dauphin is false, Elianor may know or not know. Her behavior shifts accordingly. If she knows, she is a pragmatist; if she does not, she is a desperate grandmother.
  \item \textbf{Valdais's desperation:} Valdais does not know the truth. He suspects his mother is hiding something, but he cannot prove it. He will believe whatever the party tells him—if they seem convincing.
  \item \textbf{The Cantor's role:} Alain can reveal the truth, but only if the party earns his trust. He carries the cracked ring, and he knows the name of the child Sir Corin died to save. He may not know the child is the Dauphin. He may know and have chosen silence.
  \item \textbf{The Hollow's interest:} The Ninth Taboo is a constant presence. If the party cites the ninth citation, the dead king's laughter may reveal a secret they were not meant to hear. The secret is always about the succession.
\end{itemize}

\index{Succession Crisis!Vhasia}
\index{Dauphin of Vhasia!legitimacy}
\index{Queen-Mother Elianor!succession}
\index{Valdais Longblood!claim}
\index{Cantor Alain!succession}
\index{Ninth Taboo!Vhasia}