\section{Vilikari --- ``Laurels \& Longhouses''}
\label{chap:vilikari}

\noindent\textit{The Vilikari are mercantile: traders and smugglers, but not accountants.}

\subsection*{Quick Reference}

\textbf{Tagline:} \textit{Two laws, one people. The oak remembers; the marble writes. The stone is cracking.}

\begin{quote}
  \textbf{Federate King Aldric (Elite):}  
  ``We are the bridge between empire and wilderness, the marriage of law and custom. Our strength lies not in choosing one path but in walking both with equal surety. But the bridge is rotting, and no one sends hostages anymore. The stone cracks, and the cracks have no witnesses.''
\end{quote}

\begin{quote}
  \textbf{Hearth-Mother Vesna (Commoner):}  
  ``The old ways and new laws both have their place at our table. Serve the right mead to the right guest, honor the proper customs, and you'll find we're better hosts than most kings. The trouble is, the guests are forgetting why they came. The stone is forgetting their names.''
\end{quote}

\noindent\textbf{Genre / Mood:} Early Holy Roman Empire on the steppe --- legal dualism, hybrid identity, frontier pragmatism, forgotten treaties.

\noindent\textbf{Starting Location:} A longhouse quarter tucked into an old forum. Laurel garlands hang from broken columns. Smoke from a dozen hearths curls under cracked marble arches. A Hearth‑Mother offers a cup of small beer and a piece of oatcake. ``The Foedus Stone cracks,'' she says. ``The steppe listens. Which law will you speak when the Ykrul envoy arrives --- oak or marble? The stone has a new crack. No one knows which promise it represents. Speak carefully. The stone is listening too.''

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Vilikari are \emph{forgotten}. They do not choose between empire and steppe; they marry them, awkwardly, imperfectly, and with frequent arguments. Their halls are longhouses raised over ruined forums; their laws carved on both oak and marble; their loyalties braided like the war-maidens' hair. They read the geometry of \emph{jurisdiction}. They are the bulk of the non-elite population in Ecktoria---the legionaries' descendants who never went home. A Vilikari blacksmith in Ecktoria might swear by the marble bench; his cousin two valleys over in Acasia still ties knots for the oak. This is not contradiction. It is survival.

But survival requires witnesses, and the witnesses are dying. The Foedus Stone, ancient guarantor of their autonomy, is \emph{cracking}. Every season, another shard falls. The Ecktorian clerks have stopped renewing the oath; the Ykrul no longer send hostages. The Vilikari are becoming invisible to both sides. Their law still holds, but the witnesses are fading. The Elder of Elders speaks once a year, but fewer and fewer listen. The newest crack appeared last winter. It has no date. No one knows which promise it represents.

On the coast, fishing villages double as smugglers' coves, running fugitives, forbidden goods, and whispered messages. Their boats are fast, their pilots know every reef and patrol route, and their loyalty is to the cargo's survival. They are the last people in the Amaranthine who still remember the old treaty's third clause. They are not sure why it matters. They keep it anyway.

\vspace{2mm}
\textbf{What you notice first:} Woodsmoke and old parchment. The way everyone speaks two languages at once---switching mid-sentence. Longhouse roofs that use broken imperial tiles as shingles. The silence where there used to be oaths.

\textbf{The one rule that kills newcomers:} Never invoke only one law. The \emph{mallus} (oak‑ring) and the \emph{basilica} (marble bench) must both be honored. Choose one, and the other will find you. But now, neither may answer. The stone has forgotten which law applies---so it applies neither.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Crack}

Here, the past is a weapon and the future is a weight — the weight of a stone that remembers every oath and every betrayal. The Vilikari live in the shadow of two legal systems that no longer trust each other, and in the growing silence of a witness that is slowly forgetting. Every negotiation is a tightrope between oak and marble; every handshake is a gamble that the other party will remember which law they swore by. The Foedus Stone is not a monument; it is a timer, and its cracks mark the hours until the Vilikari cease to exist as a recognized people. On the coast, the smugglers laugh at treaties and live by the tide; in the hill forts, the old legionaries drill with ghosts; in the longhouses, hearth-mothers pour beer for guests who no longer know why they came. The stone listens. The stone waits.

\begin{tcolorbox}[colback=black!3,colframe=red!60!black,title={The Ninth Taboo Manifestation}]
In Vilikari courts, the ninth witness called to a \textit{mallus} (oak‑ring) is always a ghost. The wise stop at eight and settle out of court. Those who call a ninth witness find that the ghost speaks the truth — but the truth is always the worst possible version, and the living witnesses remember it as their own. The stone's cracks are full of these truths.
\end{tcolorbox}

\subsection*{Names of the Vilikari}
\label{sec:vilikari-names}

Vilikari names blend Utaran (Latin) roots with steppe influences: hard consonants, fluid vowels, and surnames that recall a ford, a bridge, or a longhouse.

\begin{longtable}{|p{0.25\textwidth}|p{0.25\textwidth}|p{0.2\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Aldric & Vesna & de la Marche & \textit{the Two-Law} \\
\hline
Rurik & Theodora & Longwood & \textit{the Fox} \\
\hline
Drago & Gisela & Kalderash & \textit{Bridge-Born} \\
\hline
Oswin & Mira & Oakenshield & \textit{the Unbuckled} \\
\hline
Vidar & Svanhild & Tregard & \textit{Ash-Prow} \\
\hline
Kael & Ragna & Foedus & \textit{the Grey} \\
\hline
\end{longtable}

%-------------------------------------------------------------------------------
% CARD LISTS
%-------------------------------------------------------------------------------
\newcounter{cardrank}
\newcommand{\card}[4]{%
  \par\smallskip%
  \noindent\textbf{#1. #2}\hfill\textsc{\small #3}%
  \par%
  \noindent\textit{#4}%
  \par\smallskip%
}

\subsection*{Quick use notes}
Draw until you have all four suits: \textbf{Spade} = place, \textbf{Heart} = actor, \textbf{Club} = pressure, \textbf{Diamond} = leverage. Highest rank sets the main Clock: 2–5 → 4 segments, 6–10 → 6 segments, J/Q/K → 8 segments, A → 10 segments.

\textbf{Diamonds} are codified outcomes (charters, rights, escorts) that change position rather than call for a roll.

If any \textbf{Ace} appears, echo \textbf{law and bond} motifs—oaths that bind, stones that witness, treaties that remember every hand—but also \textbf{cracks and forgetting}. The stone's newest crack widens; the Hollow listens.

\subsection{Spades — Places}
\label{sec:vilikari-places}
\noindent\textit{March Town, Villa-Fort, Old Road}

\card{2}{Longhouse Quarter}{[Community]}{The forum's marble floor is worn smooth by dancing feet, not sandals. The ghosts of tax collectors weep in the corners.}

\card{3}{Milefort XVII}{[Military]}{Earth bank, stone gate; faded eagle over fresh knotwork. The eagle's stone wings were reglued after a lightning strike. It still leans left — toward the steppe.}

\card{4}{Stone Ford}{[Trade]}{Plank-palings and a customs table at knee-height; border as bureaucracy. The ford changes hands every season. The customs table never moves. The wood is stained with ink and blood.}

\card{5}{Villa Granary}{[Military]}{Barracks with a tile roof; spear racks where olives once slept. The olive press still works. The pressers are now armorers. The oil is for lamps, not cooking.}

\card{6}{Twin Court}{[Law]}{Mallus ring beneath an oak, basilica bench across the lane. The oak and the bench are the same distance from the well. The well is neutral ground. The well has a chain.}

\card{7}{Blackwood Road}{[Travel]}{Milestones re-carved in runes and Latin both. The runes say ``safe passage.'' The Latin says ``toll payable twice.'' The third line is scratched out.}

\card{8}{Staple Market}{[Trade] [Social]}{Frontier where Ykrul horses meet Vilikari iron. The horse-traders and smiths speak a pidgin of gestures and grudges. The scales are honest — mostly.}

\card{9}{Burial Field}{[Death]}{Ship-ridge mounds beside tiled mausolea, both garlanded. The mausolea have fresh offerings. The ancestors are not picky. The ship-mounds whisper at dawn.}

\card{10}{Repaired Bridge}{[Travel] [Law]}{An arch of stone, a Vilikari trestle stitched on; past and present joined. The trestle creaks in Vilikari, the stone groans in Utaran. They argue at midnight. The toll-keeper listens.}

\card{J}{Hill Amphitheater}{[Authority]}{A lord's hall raised in an old arena's curve; power in architecture. The lord's throne is a gladiator's chair. He does not know who last sat in it. The sand beneath is still dark.}

\card{Q}{New Raivon}{[Urban]}{March capital; grid-streets colliding with wandering lanes. The grid was laid by a surveyor who never came back. The lanes are older. They move at night.}

\card{K}{Dux's Palace}{[Authority]}{Purple awnings, wolf-hides, law tablets on the wall. The tablets list every treaty broken. The wolf-hides are from the treaty-breakers. They still growl.}

\card{A}{Foedus Stone}{[Law] [Legend]}{At a crossroads; every treaty carved, every hand weighed. The stone has grown. Each new treaty adds a finger of granite. The oldest fingers are worn smooth. The newest crack has no date.}

\subsection{Hearts — People \& Factions}
\label{sec:vilikari-people}
\noindent\textit{Federate, Mixed Court, Diaspora}

\card{2}{Hearth-Mother Vesna}{[Social]}{Keeper of guest-right. She has never lost a quarrel; her voice is soft, her memory longer than the Foedus Stone. She remembers obligations that are not yet owed—and the stone is forgetting, shrinking.}

\card{3}{Shield-Brother Rurik}{[Military] [Oath]}{Comitatus oath-bearer. His braids are his record; each knot is an oath kept — or broken. He has started tying knots in his sleep, hoping the stone will remember.}

\card{4}{March Notary Theodora}{[Law]}{Utaran-trained legalist. Her left hand writes Latin, her right writes runes. She signs with her foot. A ``third hand'' writes in ink that no one can read—not even her. The document is always binding—until the stone forgets.}

\card{5}{Horse-Reeve}{[Trade]}{Rations pasture between plows and warbands. He measures grass in fistfuls. The Ykrul respect his arithmetic. The Ecktorians pretend not to notice. He is saving fistfuls to patch the stone.}

\card{6}{Ykrul Envoy}{[Social]}{Riding under safe-conduct; speaks obligations in proverbs. He never carries a weapon. His tongue is sharp enough. His hostage string has nine knots. He is running out of space—and the stone is running out of memory.}

\card{7}{Old Legionary}{[Military]}{Drill-master for Vilikari levies; Utar in muscle memory. He still marks time in paces. The recruits call him ``Grandfather Eagle.'' The eagle on his shield is chipped. He says the stone's cracks are just more places to put his thumb.}

\card{8}{Ecktorial Factor}{[Trade]}{Vilikari blood, Ecktorial coin; flows where the market points. His coat has two collars—one for each loyalty. He wears whichever is warmer. His tally has three columns. The third column is labelled ``crack width.''}

\card{9}{Kahfagian Pilot}{[Travel]}{Lantern-law chit and clan tattoo; sea-law on land. He navigates by stars and tariffs. Both change with the season. He has never been lost—only delayed. The stone's cracks are good for smuggling.}

\card{10}{Marcher-Lord}{[Authority]}{Iron-nail crown, Vilikari law in pocket; a frontier king. His crown was a plowshare. He reversed the curse. His hands are still calloused. He does not sit on his throne. He spends his days pressing his ear to the stone.}

\card{J}{War-Maiden}{[Combat] [Oath]}{Oath-braid names the dead and the owed; vengeance as memory. Her braid is longer than her sword. Both have drawn blood. She never cuts the braid. The dead would forget her. The stone is forgetting everyone else.}

\card{Q}{Queen of Marches}{[Politics]}{Consort-regent; smiles like precedent. She married the Federate King for peace. She kept the kingdom for herself. Her seal is a broken arrow. She keeps a piece of the stone in her pocket—it gets smaller every year.}

\card{K}{Federate King Aldric}{[Authority]}{Rex-Dux sworn to forms, sung by his own. He has three thrones—oak, marble, and saddle. He uses whichever is not being argued over. Lately, all three are dusty. He has stopped arguing. He just listens to the stone.}

\card{A}{Elder of Elders}{[Law]}{Thing-council speaker; a nod can unmake a clan. He speaks once a year. The rest of the time he listens—and remembers. But his memory is cracking like the stone. He has started writing things on his own skin.}

\subsubsection*{Other Prominent Figures}

\noindent\textbf{Smuggler Captain Drago:} Vilikari fisherman and trafficker. He runs contraband and refugees past the Kahfagian patrols, knows every reef and every lantern shift, and claims the Foedus Stone's cracks are good for business—no one checks the old treaties anymore.

\subsection{Clubs — Complications \& Threats}
\label{sec:vilikari-complications}
\noindent\textit{Two Laws, Two Fronts, Old Grudges}

\card{2}{Annona Late}{[Supply] [Social]}{Grain stipend misses the comitatus; tempers quicken like flint. The granary is full. The keys are in the wrong vault. Someone is waiting for the price to rise. The soldiers are not. The stone does not care.}

\card{3}{Law Tangle}{[Law]}{Mallus verdict vs. basilica writ — whose seal bites? The oak judge and the marble judge are cousins. They send each other holiday cards. The case has been open for three generations. The stone has forgotten who started it.}

\card{4}{Raid Rumor}{[Combat] [Fear]}{A Ykrul boyar; levies panic the plowlands. The boyar is a ghost—or a tax collector. The farmers don't wait to find out. They burn the fields themselves. The stone's cracks are good cover for ghosts.}

\card{5}{Bridge Levy}{[Trade] [Combat]}{Doubled for road-works; caravans balk, fists rise. The road-works are a ditch. The ditch fills every spring. The bridge keeper has a new sword. He sharpens it on the stone.}

\card{6}{Succession Feud}{[Politics] [Death]}{Wolf-brother claim vs. purple-charter heir. The wolf-brother has a better sword. The purple-charter has a better notary. The notary's ink is mixed with blood. The stone's crack drinks it.}

\card{7}{Foedus Recall}{[Law] [Social]}{A prefecture resettlement is ``requested''; homes bristle. The prefecture's map is forty years old. It shows a village that drowned. The villagers are still there. The stone remembers their names—barely.}

\card{8}{Coin Talk}{[Trade] [Weight]}{Debasement; pay demanded in salt, hides, or hostages. The salt is from a disputed pan. The hides are from cattle that were never born. The hostages are children. The stone's cracks are full of their names.}

\card{9}{Grave Offense}{[Death] [Omen]}{A burial field disturbed; omens ride with the news. The ancestors are angry. They have a list. Your name may be on it. The list is carved on a ship's prow. The stone's new crack is shaped like a finger pointing.}

\card{10}{Port Clash}{[Law] [Trade]}{Lantern-law vs. street-law; confiscations bloom. The lanterns were moved last night. The old safe channel is now a reef. The harbor master is not answering questions. He is listening to the stone.}

\card{J}{Warband Flip}{[Military] [Betrayal]}{Colors change mid-pay; hostage protocol snaps shut. The hostages are eating well. Their captors are not. The captain's new banner is stitched with his old captain's hair. The stone's crack has grown wider.}

\card{Q}{Restoration Preacher}{[Social] [Fear]}{Stirs crowds against the ``barbarian king in purple.'' The preacher's license was signed by a dead abbot. The seal is still warm. The crowd does not know. The stone does not care.}

\card{K}{General Levy}{[Military] [Supply]}{The march is closed; markets freeze into spears. The levy notice is posted in three languages. None of them are local. The tax collector has a bow. The stone has stopped listening.}

\card{A}{Spring Melt}{[Weather] [Travel]}{Eats the Wolf Road; proofs, plans, and armies bog. The road is a memory. The ruts fill with water—and older things. The bridge keeper is fishing for boots. The stone's crack is now a chasm.}

\subsection{Diamonds — Rewards \& Leverage}
\label{sec:vilikari-rewards}
\noindent\textit{Charter, Right, Escort}

\card{2}{Foedus Seal}{[Law]}{Settlement and stipends confirmed for a season. The seal is a bronze wolf's head. It warms when the treaty holds—and cools when it doesn't. The wolf's eyes are garnets. One eye is cracked.}

\card{3}{Mallus Right}{[Law]}{Hold court under the oak and be obeyed. The oak is older than the kingdom. Its roots have seen more verdicts than the throne. The acorns are bitter. The oak's leaves are falling early.}

\card{4}{Stipend Arrears}{[Supply]}{Paid in salt and grain; portable, persuasive. The salt is grey. The grain is from last harvest. Both are worth more than coin. The soldiers will accept. The stone will not.}

\card{5}{Utaran Patent}{[Social]}{Citizenship-style privilege; doors open, bows deepen. The patent is a worn scrap of vellum. It has been folded and refolded for a century. The seal is a sunburst. The sunburst is fading.}

\card{6}{Stall-Right}{[Trade]}{Customs favor on a frontier market line. The stall is a painted plank. Plant it, and the tax man will pretend not to see you. He will remember the favour. The stone will not.}

\card{7}{Hostage Treaty}{[Social] [Oath]}{Exchange fosters; grudges trade for leverage. The hostages are children. Their parents are patient. The treaty has nine clauses. The ninth is unspoken. The stone has forgotten the first eight.}

\card{8}{Remount Allotment}{[Travel]}{Fresh horses from royal paddocks, no questions. The horses are from the king's own herd. They know the roads better than you. They also know the king's secrets. They refuse to cross the stone's shadow.}

\card{9}{Wergild Table}{[Law] [Weight]}{Recognized by both codes; vengeance priced and paid. The table is a slate slab. The prices are carved deep—and often recarved. The newest price is a year of rain. The rain is washing away the carvings.}

\card{10}{Bridge Farm}{[Trade]}{Imperial for a season; funds and friends accrue. The bridge toll is a copper a head. The bridge keeper is your new best friend. He also keeps a list of your enemies. He scratches the list on the stone.}

\card{J}{Purple Warrant}{[Authority]}{Wear the stripe; officials defer on sight. The stripe is a sash. It was a curtain in a past life. The present life is yours. Do not lose it. The stone's crack has swallowed one like it.}

\card{Q}{Queen's Audience}{[Social]}{Private; a whisper becomes policy. The audience is held in a garden. The flowers are listening. The gardener is a spy. The gardener is also the queen. She keeps a piece of the stone under her tongue.}

\card{K}{Dux Commission}{[Military]}{Command a mixed unit with pay authority. The commission is a baton. It doubles as a mace in a pinch. It also doubles as a receipt. The receipt is overdue. The stone has a receipt for the stone.}

\card{A}{Great Law Day}{[Law] [Legend]}{Proclaim a new concord; one temporary exception to old law. The proclamation is made at noon. By sunset, everyone will have an opinion. By dawn, the Foedus Stone will have a new finger—or a new crack.}

%-------------------------------------------------------------------------------
% MECHANICS
%-------------------------------------------------------------------------------
\subsection{Vilikari Mechanics}
\label{sec:vilikari-mechanics}

\subsubsection{Core Clocks}

\paragraph{Witness Fading Clock [8]}
Tracks the erosion of Vilikari legal authority.

\textbf{Ticks when:}
\begin{itemize}[nosep]
  \item The party defies a Vilikari judgment or brings a dispute to an outside court (+1).
  \item The GM spends 2 SB to represent a forgotten oath (+1).
  \item A foreign envoy (Ykrul, Ecktorian, etc.) publicly ignores a Vilikari ruling (+1).
\end{itemize}

\textbf{At 4 segments:} Any Vilikari NPC the party interacts with suffers –1 die to social rolls (they doubt their own authority). The stone's cracks are visible to all.

\textbf{At 8 segments:} The Foedus Stone cracks completely. The Vilikari are no longer recognized as a neutral party. The Ykrul and Ecktorians begin fighting over their land. The party can reset the timer only by \textbf{renewing the treaty}—finding a witness older than the stone itself (a Lethai-al elder, a Trow, a ghost of the original signatory) and having them speak the oath aloud.

\paragraph{Two-Law Tension [4]}
Whenever a \emph{jurisdictional conflict} arises (mallus vs. basilica, oak vs. marble), players may step Position up by citing the Foedus Stone or an older precedent—but every use adds a tick to the Two-Law timer. When it fills, one side demands exclusive authority. The other side sharpens its knives.

If the Witness Fading Clock is at 4+, this timer ticks twice as fast.

\paragraph{Border Tension [Front]}
Once per session, if a deal is struck with a foreign envoy or mixed-blood courtier, advance or start this front.

\paragraph{Market Knives [4]}
At any \emph{Staple Market} or \emph{Stone Ford}, each time a player converts a held Diamond into a \emph{Market Edge} (+1 die to all buy/sell rolls for the scene), tick this timer. When it fills, a rival factor challenges your right to trade—face a duel, a bribe demand, or a confiscation.

\paragraph{Feast-Bond [3] / Blood-Owed [3]}
Created by the Treaty-Feast Procedure. While Feast-Bond has segments, both parties are protected from betrayal in that matter. On breach, the breaker immediately takes 1–2 SB and the timer flips to Blood-Owed [3].

\subsubsection{Status \& Environment}

\paragraph{The Two-Law Dial (Oak $\leftrightarrow$ Marble)}
Track which legal tradition currently holds sway.
\begin{itemize}[nosep]
  \item \textbf{Oak High:} +1 Position when invoking mallus, blood-price, or tribal custom; the first basilica petition starts one step worse.
  \item \textbf{Marble High:} +1 Effect when citing charters, patents, or Utaran precedent; the first mallus oath starts one step worse.
\end{itemize}
Center the dial by offering a wergild recognized by both sides or hosting a joint moot. If the Witness Fading Clock is at 4+, centering requires a sacrifice (a memory, a name, or a drop of blood on the stone).

\paragraph{Frontier Diplomacy}
If a Heart card involves a foreign envoy, a federate, or a mixed-blood courtier, treat the first Diamond played in the scene as doubled leverage. If the Witness Fading Clock is at 6+, the deal also ticks the timer by 1 (the stone forgets another clause).

\paragraph{March Markets}
At any \emph{Staple Market}, \emph{Stone Ford}, or \emph{Repaired Bridge}, Diamonds tied to rights or escorts may be cashed as currency in lieu of coin, one-for-one. Doing so creates a \textbf{Stone Weight} that the GM may call in later through factions or raiders. The weight is never written. It is remembered—until the stone forgets.

\paragraph{The Vilikari Diaspora}
Vilikari characters in Ecktoria gain +1 die on Streetwise checks (they know the back alleys and the bribes), but –1 die on social rolls with Old Blood patricians (who see them as upstarts). In Acasia, they gain +1 die on Survival checks (their ancestors were farmers here) but –1 die on social rolls with the Curse-touched (who remember imperial tax collectors). A Vilikari character may start with one ``family tally''—a minor asset that tracks a single unpaid favour across generations. If the Witness Fading Clock ever fills, the tally turns to ash.

\subsubsection{Procedures}

\paragraph{Treaty-Feast Procedure (Fast Diplomacy)}
\begin{enumerate}[nosep]
  \item \textbf{Set the Table:} Name 2–3 \emph{terms} (what both sides want) and 1 \emph{taboo} (off-limits).
  \item \textbf{First Cup (Position):} Roll a social test; success grants +1 die to bargain, failure grants the host 1 SB.
  \item \textbf{Bread \& Salt (Offer):} Put one leverage on the table (Diamond or oath-promise). The host matches with an obligation (escort, market access, or hostage).
  \item \textbf{Smoke (Complication):} The GM reveals a pressure tied to \emph{Old Grudges} or \emph{Two Laws}. Resolve it or accept a \emph{Rider Clause}.
  \item \textbf{Knot the Braid:} On success, create a Feast-Bond [3] timer. On breach, the breaker takes 1–2 SB and the timer flips to Blood-Owed [3].
  \item \textbf{Stone Witness:} If the Feast-Bond is sworn at the Foedus Stone, reduce the Witness Fading Clock by 1. The stone remembers one more promise.
\end{enumerate}

\paragraph{Frontier Market (Quick Trading)}
At any Staple Market or Stone Ford, players may convert a held Diamond into a Market Edge. Each conversion ticks Market Knives [4]. When the timer fills, a rival factor challenges your right to trade.

If the Witness Fading Clock is at 6+, the rival factor is a ghost—a forgotten signatory of the Foedus Stone. The duel is a legal dream-duel.

\paragraph{Start-of-Scene Omens}
Roll \textbf{1d6} at the start of a Vilikari scene:

\begin{enumerate}[nosep]
  \item \textbf{Laurel Wilt:} First legal action gains +1d; first boast draws \textbf{Law Tangle}.
  \item \textbf{Smoke on the Steppe:} \textbf{Raid Rumor} advances once; shortcuts open.
  \item \textbf{Twin Bells:} Both courts are in session; pick a side or start \textbf{Law Tangle}.
  \item \textbf{Queen's Dove:} A message arrives; +1d on the next \textbf{Queen's Audience}.
  \item \textbf{Grey Salt:} \textbf{Stipend Arrears} counts double value today; smugglers' ears open.
  \item \textbf{Old Eagle:} An Utaran relic surfaces; \textbf{Utaran Patent} use is free this scene.
\end{enumerate}

\noindent\textbf{Crack Light (special):} If the Witness Fading Clock is $\geq$ 4, the stone also glows. The next legal argument made at the stone is automatically binding—for both sides.

\subsubsection{Superstitions (burn for +1 die once)}

\begin{itemize}[nosep]
  \item \textbf{Bread to the Stone:} Crumble a crust at the base of the Foedus Stone. Ignore the first \emph{Law Tangle} penalty.
  \item \textbf{Knot the Braid:} Tie a single knot in a horse's tail. Cancel \emph{Annona Late} or take –1 die on your next supply roll (your choice).
  \item \textbf{Name to the Oak:} Whisper a dead enemy's name into the hollow of the mallus oak. Gain +1 Effect when collecting wergild from their kin this scene, but mark \emph{Obligation} to the Elder of Elders.
  \item \textbf{Press the Crack:} Touch the newest crack in the Foedus Stone and speak a promise you intend to break. Gain +1 die on your next deception roll, but the stone will remember your name. The Hollow will too.
\end{itemize}

\subsubsection{Strain Budget (SB) Spends}

\paragraph{Minor (1 SB)}
\begin{itemize}[nosep]
  \item \textbf{Exposure:} March officials or tribal elders take notice.
  \item \textbf{Noise:} Levy guards or market patrols are alerted.
  \item \textbf{Trace:} Trackers or treaty keepers mark your route.
  \item \textbf{Delay:} Lose a favorable legal window or crucial time.
  \item \textbf{Supply Strain:} Mark +1 segment on a relevant resource timer.
  \item \textbf{Crack Whispers (Vilikari only):} A character hears the stone whispering a forgotten promise. Test Resolve (DV 3) or mark 1 Fatigue.
\end{itemize}

\paragraph{Moderate (2 SB)}
\begin{itemize}[nosep]
  \item \textbf{Alarm Raised:} The Horse-Reeve or March Notary responds.
  \item \textbf{Position Lost:} Legal challenge or border closure costs you advantage.
  \item \textbf{Foe Appears:} A rival faction or warband arrives.
  \item \textbf{Gear Trouble:} A key item becomes Compromised or Neglected.
  \item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
  \item \textbf{Witness Fading:} Advance the Witness Fading Clock by 1 segment.
\end{itemize}

\paragraph{Serious (3 SB)}
\begin{itemize}[nosep]
  \item \textbf{Reinforcements:} Shield-brothers, levy forces, or foreign troops arrive.
  \item \textbf{Key Gear Breaks:} A crucial tool or weapon becomes temporarily unusable.
  \item \textbf{Major Twist:} Treaty voided, levy called, or border dispute erupts.
  \item \textbf{Rail Tick:} Advance a relevant campaign/front timer by 1 segment.
  \item \textbf{Condition Applied:} Mark Fatigue 1, Harm 1, or an appropriate Condition.
  \item \textbf{Stone Speaks:} The Foedus Stone whispers a forgotten clause that invalidates a character's legal protection for the scene.
\end{itemize}

\paragraph{Major (4+ SB)}
\begin{itemize}[nosep]
  \item \textbf{Trap Springs:} A prepared danger activates with full effect.
  \item \textbf{Authority Arrival:} The Federate King, Queen of Marches, or Elder of Elders intervenes.
  \item \textbf{Scene Shift:} A raid hits, the levy is called, or a treaty breaks.
  \item \textbf{Patron Omen:} Divine or arcane forces manifest—an omen appears, a blessing is lost, or a curse manifests.
  \item \textbf{Narrative Pivot:} The story reframes your objectives unexpectedly.
  \item \textbf{Crack Splits:} The Foedus Stone's newest crack widens. One of the party's active Diamonds crumbles to dust.
\end{itemize}

\paragraph{Region-Specific SB Conversions}
\begin{itemize}[nosep]
  \item \textbf{Seal Crack (1 SB):} A writ loses half its value until re-witnessed.
  \item \textbf{Citation War (2 SB):} Increase DV by 1 unless you produce a precedent (spend a Diamond or recall a past ruling).
  \item \textbf{Crown Fracture (3+ SB):} A treaty splits; halve a consequence by accepting a rival's seal on your paper (start or tick Two-Law timer).
  \item \textbf{Stone Whisper (2 SB, Vilikari only):} The Foedus Stone reveals a forgotten clause. The GM must answer one yes/no question about a treaty or oath in the region.
\end{itemize}

\subsubsection{Reputation Echoes}

\begin{itemize}[nosep]
  \item \textbf{Court-Favored:} Once per session, downgrade \textbf{Law Tangle} to a minor delay; the next document you present is scrutinized (+1 DV).
  \item \textbf{Raid-Tested:} +1 Position when evading Ykrul or Black Banners; supply lines run 1 segment longer.
  \item \textbf{Neutral Broker:} +1d with both oak and marble in a treaty scene; any betrayal earns \textbf{Blood-Owed} [2] from both sides.
  \item \textbf{Stone-Marked:} Gained by renewing the treaty. The Foedus Stone remembers your name. Once per session, reroll a failed social roll involving a Vilikari or a treaty. The stone will ask for a memory in return.
\end{itemize}

\subsubsection{Local Patrons}

\begin{itemize}[nosep]
  \item \textbf{The Two-Faced King} — Mykkiel. \textit{Justice that wears a laurel and a wolf's tail.}
  \item \textbf{The Hearth Judge} — The Witness. \textit{The flame that watches both sides of the bargain.}
  \item \textbf{The Oak Father} — Grimmir. \textit{The root that holds the law; the branch that signs the treaty.}
  \item \textbf{The Coin Saint} — Maelstraeus. \textit{Interest on a handshake; the stone that never sleeps.}
\end{itemize}

\subsubsection{Boss Seeds (Three-Phase Foes)}

\paragraph{The March Notary (quills like daggers)}
\emph{Phases:} Citation Tangle $\rightarrow$ Seal Swap $\rightarrow$ Precedent Trap.\\
\emph{Cracks:} Produce a \textbf{Foedus Seal} to quiet him; a \textbf{Queen's Audience} overrules his filing. If the Witness Fading Clock is at 6+, he gains +1 die in all phases (the stone supports him).

\paragraph{Ykrul Storm-Rider (wolf-cloak, bone whistle)}
\emph{Phases:} Raid Cry $\rightarrow$ Herd Stampede $\rightarrow$ Blood Price.\\
\emph{Cracks:} A \textbf{Wergild Table} settles the feud without combat; a \textbf{Hostage Treaty} buys a season's peace. He fights harder if the stone is cracked (he smells weakness).

\paragraph{The Prefect's Auditor (Ecktorial ink on his fingers)}
\emph{Phases:} Tally Seizure $\rightarrow$ Tax Levy $\rightarrow$ Fort Closure.\\
\emph{Cracks:} An \textbf{Utaran Patent} grants immunity; a \textbf{Mallus Right} transfers jurisdiction to the oak. He carries a piece of the stone in his pocket; if the party takes it, the Witness Fading Clock resets by 2.

\paragraph{The Stone's Ghost (Foedus Avatar)}
\emph{Phases:} Whisper of Forgotten Clauses $\rightarrow$ Crack Widens $\rightarrow$ The Hollow Steps Through.\\
\emph{Cracks:} Speak the original treaty aloud (requires a Lethai-al witness), repair the stone with a memory of a kept promise, or sacrifice a personal oath to feed the stone instead of the Hollow.

%-------------------------------------------------------------------------------
% LORE & CROSS-REGION
%-------------------------------------------------------------------------------
\subsection{Lore Echoes}
\label{sec:vilikari-lore}

\begin{itemize}[nosep]
  \item \textbf{The Smugglers' Covenant (The Sealed Gate):} Vilikari pilots who run the Kahfagian coast swear by a secret oath to the \emph{Unlatched Lock}—a threshold that never fully closes. They leave a copper nail at every hidden cove. A captain who breaks the covenant finds that their boat will not cross the tide line; the sea simply refuses to carry them. The Kahfagian Admiralty has tried to bribe the covenant's keeper; he accepted the coin and sank the next morning in calm water. The stone's newest crack appeared that same night.
  \item \textbf{Saint Theodora of the Dual Writ (The Witness):} A march notary who uncovered a Fenwood spy in the Federate King's court. She forced the spy to sign a confession under both oak and marble, then burned the oak copy and kept the marble one. The spy vanished, but his signature appears on every broken treaty in Vilikari history. Her quill is kept in the Dux's Palace; it writes in lemon juice, and the truth it records cannot be erased by fire or water. The stone's cracks drink the juice.
  \item \textbf{The Host of the Two Ledgers (The Longhouse Healers):} A hospice near the Foedus Stone is run by grey-robed healers who ask no payment. They treat wounded raiders from both sides of the steppe, and the Ykrul and Ecktorian captains have agreed to a truce around its doors. The Federate King tolerates them because they also hold the rite of the \emph{Unbuckled Oath}—a ceremony where a knight may renounce a blood-debt without losing honour, by placing their sword on the stone and speaking a single true regret. The stone remembers these regrets. It has many.
  \item \textbf{The Ostrikari Cousins (The Savage Heart):} The eastern Vilikari, called Ostrikari by lowlanders, are no more savage than any frontier people. Their reputation comes from the bear cults of the upper Wolf Road, where initiates wear cured skins and speak in growls. The Ostrikari do not raid Vilikari villages; they trade furs for iron and send their young to serve as mercenaries in the Black Banners. A Vilikari proverb says: ``The Ostrikari sharpen their teeth on stone, not on kin.'' The stone does not mind. It has no teeth left.
\end{itemize}

\subsection{Cross-Region Hooks}

\begin{itemize}[nosep]
  \item \textbf{Kahfagia (Lantern-Law):} A \textbf{Foedus Seal} may stand in for a \emph{Safe-Conduct} along a beaconed lane once. Doing so starts \textbf{Port Clash} [2] in the next harbor touched.
  \item \textbf{Ecktoria (Imperial Remnants):} An \textbf{Utaran Patent} counts as either \emph{Archive Hour} (scholars open doors) or \emph{Icon License} (display symbols) in Ecktorial sites—choose once per session.
  \item \textbf{Aelerian Mountains (Pass \& Toll):} A \textbf{Pass-Ring} (from the Aeler expansion) grants priority at dwarf switchbacks; cashing it to bypass inspection spawns \textbf{Stone Notice} [3].
  \item \textbf{Direwood (Breath \& Bells):} Spending \textbf{Stipend Arrears} for levee labor converts one Direwood \textbf{Clubs} draw into a \emph{Ward-Salt} Diamond for a scene, then begins \textbf{Names Disputed} [3] with bell-wardens.
  \item \textbf{Acasia (Curse \& Tithe):} A \textbf{Wergild Table} accepted by a Curse-touched village also clears one \emph{Feud Spark} timer, but the Hollow gains +1 Attention.
  \item \textbf{Lethai-al Silk Bridges:} A \textbf{Hostage Treaty} can be witnessed by a Lethai-al silk-warden, making it binding across the Crimson Valley. The price is a strand of the treaty-bearer's hair.
\end{itemize}

%-------------------------------------------------------------------------------
% GM GUIDANCE
%-------------------------------------------------------------------------------
\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In Vilikari society, power flows through the careful balance of multiple legal traditions, the management of frontier diplomacy, and the maintenance of complex oaths and treaties. The Federate King maintains authority through a combination of imperial forms and traditional customs, while local leaders wield influence through their knowledge of both Utaran law and tribal custom. The true power lies with those who can navigate this complex landscape while maintaining the loyalty of their people and the respect of their neighbors.

But the stone is cracking. The witnesses are dying. The Vilikari are becoming \emph{invisible}.

\textbf{For the GM:}  
Patronage in Vilikari society revolves around legal documents, diplomatic relationships, and the ability to command respect from institutions that operate under multiple legal systems. Rewards often take the form of treaties, charters, and official positions that can be leveraged into greater influence. To emphasize the hybrid identity and the fading authority:
\begin{itemize}[nosep]
  \item Tie rewards to visible symbols (seals, writs, treaties) that can be challenged, stolen, or voided—and that may crumble if the stone forgets.
  \item Let rival legal systems issue conflicting authorities, forcing players to choose whose favour matters more. Neither may remember them tomorrow.
  \item Use the courts, markets, and frontier posts as arenas for social contests, where knowledge of law and custom determines success—and where the stone's whispers can tip the balance.
  \item Remember that Vilikari characters are at home in Ecktoria's counting-houses and Acasia's hedgerow courts alike—but they are never fully trusted in either. That suspicion is their best tool. When the stone cracks fully, that suspicion becomes exile.
\end{itemize}
In Vilikari society, your word is your bond, and your bond determines whether you rule or are ruled. But a bond is only as strong as the stone that witnesses it. And the stone is forgetting.
\end{tcolorbox}

% Index entries
\index{Vilikari|see{Laurels and Longhouses}}
\index{Theme generator}
\index{Laurels and Longhouses generator}
\index{Places!Vilikari}
\index{People!Vilikari}
\index{Complications!Vilikari}
\index{Rewards!Vilikari}
\index{Federate King (Rex-Dux)}
\index{Queen of the Marches}
\index{Elder of Elders}
\index{Dux's Palace}
\index{New Raivon}
\index{Foedus Stone}
\index{Hill Amphitheater}
\index{Milefort XVII}
\index{Twin Court}
\index{Burial Field}
\index{Staple Market}
\index{Repaired Bridge}
\index{Longhouse Quarter}
\index{Blackwood Road}
\index{Ykrul Envoy}
\index{Shield-Brother}
\index{Hearth-Mother}
\index{March Notary}
\index{War-Maiden}
\index{Horse-Reeve}
\index{Ecktorial Factor}
\index{Old Legionary}
\index{Marcher-Lord}
\index{Foedus Renewal Seal}
\index{Mallus Bench-Right}
\index{Wergild Table}
\index{Purple Warrant}
\index{Great Law Day}
\index{Dux Commission}
\index{Hostage Treaty}
\index{Smugglers' Covenant}
\index{Saint Theodora}
\index{Host of the Two Ledgers}
\index{Ostrikari}
\index{The Savage Heart}
\index{Local Patrons!Vilikari}
\index{Witness Fading Clock}
\index{Foedus Stone cracks}