% ============================================================
% Viterra --- ``The Hedge-Law Realm''
% Tudor Law vs. Hedge-Law twist: Law as bloody border warfare
% ============================================================
\section{Viterra --- ``The Hedge-Law Realm''}
\label{chap:viterra}

\subsection*{Quick Reference}
\textbf{Tagline:} \textit{Every hedge is a boundary. Every boundary is a claim. Every claim is a blade.}

\begin{quote}
  \textbf{Warrior Queen Veralyn Everblood (Elite):}  
  ``I was Fenwood's ward when I was sixteen. I took Lence's walls for him. I made the Vhasian Sun-court run red. Now I sit on a throne built of precedents—and I keep a dagger under my pillow. The hedge that is not defended is only a suggestion. The law that is not enforced is only a memory. I will have both remembered—in stone, in blood, and in the ledgers of every parish from the Belworth to the Dolmis. But sometimes... I wonder which hand shakes more when I seal a charter: the queen's, or the enforcer's.''
\end{quote}

\begin{quote}
  \textbf{Fen Reeve Cuthbert (Commoner):}  
  ``The dikes hold back more than water—they hold back the scream of Lence still echoing in the reeds. Every rod measured, every stone placed, keeps the flood from washing away not just our fields but the memory of what we were before the Fenwoods came. Across the Belworth, the Vhasians still claim our hedges are theirs. Let them try. We have lawyers who quote old charters. We have swords that remember Lence. Sometimes they are the same people. The Queen's peace is a contract; she signs it each morning with a blade in her hand. I just hope the ink holds—and that it's not mine.''
\end{quote}

\textbf{Genre / Mood:} Post-imperial consolidation, legal warfare as blood sport, border adventurism, the fragile crown haunted by fratricide.

\textbf{Starting Location:} The Hall of Dawning tilt-yard at first light. Dawn-knights drill in armor polished like mirrors—each dent a story from last week's legal duel. Beside them, advocates rehearse arguments with the same intensity, palms slick on pommels. A Queen's Justiciar slides a parchment across an oaken rail, her seal ring catching the dawn. ``Three charters, one road across the Belworth. The Vhasian claimant says the bridge is theirs. The local reeve says it's crown land. The abbey says it's a pilgrimage, not a toll. Carry the Queen's writ, break the Quiet Tolls, and make them agree—by clause or by cut. The Earls are watching, and the Queen has only just secured the throne. Fail, and you'll give her enemies the excuse they need: that she is still Fenwood's bloody handmaiden, not Viterra's queen.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
Viterra breathes with the careful rhythm of precedent and petition—but also with the raw, unhealed scar of a kingdom that threw off the Utaran yoke and is still learning to stand alone. Once a proud, independent kingdom before the legions came, Viterra never forgave the empire. It was the first to break away when the Utaran Imperium faltered, its hedge-laws and parish courts preserving a stubborn, deep-rooted identity that parchment and wax could never erase. Now, under a young and untested Warrior Queen, Viterra is trying to remember what it means to rule itself.

But the crown is not secure. Queen Veralyn Everblood took the throne only after the sudden death of her brother, King Baldwyn Brightstide, who reigned for barely two years before succumbing to a swift, mysterious illness—rumors whisper it was Fenwood's doing, to clear the path for *their* claim. His death was a wound the kingdom has not healed. The succession was contested: the earls of the northern marches, who remember when their ancestors ruled as petty kings, demanded concessions; the barons of the western borderlands, many with Vhasian blood and Vhasian gold in their pockets, whispered that a woman could not hold the crown—especially one who learned statecraft at the elbow of the duke who sacked Lence. The Queen has spent her first year in the saddle, riding from keep to keep, sealing charters with her own seal, settling feuds with her own justiciars, and—when necessary—proving her right to rule with a blade in her hand. The throne is secured, for now. But the kingdom is a patchwork of loyalties, and every hedge hides a knife.

The realm's true sinews lie not in steel but in writs, charters, and the ancient rights that bind lord to tenant, crown to county, and church to common. Every hedgerow, every parish stone, every ferry crossing carries the weight of custom older than memory. Yet there is no single, unified law. The Queen's law competes with the earls' law, with the barons' law, with the ancient custom of the fens, with the Ubral clan law that still holds in the high passes, and with the Vhasian claims that bleed across the Belworth. A deed that is valid in Valora may be worthless in the Marches. A judgment that holds in a crown court may be overturned by a hedge-moot at dusk. The very question \emph{which law is law} is the central argument of the kingdom—and it is settled not in courts, but in the mud of the Belworth, where writs are brandished like blades and blood soaks into the parchment.

Viterra is not a nation at peace. It is a realm that wages law as war, and war as law. The Queen's Peace is not a promise—it is a contract, renewed each morning by every magistrate, every justicar, every clerk who opens a ledger. And the ink is still wet on many pages—sometimes with more than just ink.

\vspace{2mm}
\textbf{What you notice first:} The smell of damp parchment laced with old iron. The clatter of writ-runners' boots on cobblestones, heels striking sparks where old blood soaks into the stone. The way every noble's hand rests on a seal ring worn thin from constant gripping—and every commoner's hand rests on a hatchet notched with Vhasian blood.

\textbf{The one rule that kills newcomers:} Never cross the Belworth without a charter recognized on both banks. Viterran law and Vhasian claim fight every crossing, and the river takes the losers—sometimes still clutching a forged writ to their chest.
\end{tcolorbox}

\subsection*{Regional Mood: Hedge, Crown, and Claim}

Here, law is a blade. The Warrior Queen rules from Valora, her throne a bench of precedents. She is young, untested, and surrounded by counselors who mistake caution for loyalty—and by courtiers who still whisper the name of her dead brother, wondering if the illness that took him was truly natural. The old Fenwood dukes—whose ancestors sacked Lence and left the Vhasian Sun-court fractured—still hold vast estates in the northern marches, and their claims across the Belworth keep the border smoking like a low fire. The earls of the north, descendants of the old hill kings, watch her every move, waiting for a misstep that will prove she is still Fenwood's instrument. The barons of the west, many of whom swore oaths to Vhasian claimants before they swore to her, smile with teeth that glitter with coin from across the river—and with the ghost of Lence in their eyes. Every spring, the Queen's justiciars ride west with warrants and escorts, demanding homage from villages that have paid Vhasian tithes for a generation. Every autumn, Vhasian routiers raid the grain convoys, and the legal duels that follow are fought with steel and citation alike. The border is not a line on a map. It is an argument. And the Queen intends to win it—even if it means becoming the thing she swore to leave behind.

North, the Ubral clans watch from their cairns, ready to rekindle feuds if a Viterran levy crosses the passes. South, Thepyrgos sends clerks and spies, testing the edges of Viterran jurisdiction with every disputed toll. And in Theona, the Three Greens mumble old curses while Viterran bailiffs try to count their cattle—and always miscount by one. The Queen's rule is not yet secure. Every charter she signs is a challenge. Every justiciar she sends is a declaration. Every duel fought in her name is a step toward a throne that is still being carved—out of doubt, out of blood, out of the quiet terror that perhaps she is not the ruler Viterra needs, but the enforcer it deserves.

\subsection*{Prominent NPCs of Viterra}
\label{sec:viterra-npcs}

\begin{longtable}{|p{0.22\textwidth}|p{0.26\textwidth}|p{0.32\textwidth}|p{0.1\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation \& Manner} & \textbf{Rumoured Quirk} \\
\hline
Queen Veralyn Everblood & Warrior Queen of Viterra & Haunted by Lence, hesitates on the throne—her hand shakes when signing charters, not from weakness, but the memory of blood on those same fingers. & Sleeps with a dagger under her pillow; touches it before every seal. Her crown has no jewels—only seals from conquered claims, each one cold to the touch. \\
\hline
Justiciar Elodie Venn & Queen's Justiciar & Raised from the fen marshes; her legal duels leave opponents not just defeated, but *ashamed*—they tend to 'find stairs' only after realizing they've wasted her time. & Has never lost a duel; carries a writ signed in Baldwyn Brightstide's blood as her lucky charm. \\
\hline
Baron Aldric & Disgraced turncloak & Stripped of lands for oath-breaking during the succession crisis; now patrols the northern marches not for pardon, but to prove he still bleeds Viterran—and to watch for the Queen's moment of doubt. & Wears his stripped baron's belt as a sash; speaks only in legal precedent when agitated. \\
\hline
Ser Roran & Knight of the Queen's household & A Vhasian prisoner the Queen bought back with silver and a promise; his gauntlet is a reminder that mercy is a choice, not a weakness. & Never takes ransom; his left gauntlet is silver, etched with the Vhasian sun he once served. \\
\hline
Fen Reeve Cuthbert & Keeper of the dike guilds & The dikes are his confessional; he measures floods not to save crops, but to atone for the Fenwoods' sins—and to ensure the Queen's tax men don't drown in the same waters that took Lence. & His tally-rod has notches from three floods; the next notch will be from a Vhasian raid he *let* happen to test the Queen's resolve. \\
\hline
Master Advocate Halric & Legal duelist & Believes law is the only true weapon left in Viterra; won't take sides in the succession dispute—which makes everyone suspect him of playing both ends. & Has argued cases in Vhasian, Thepyrgos, and Ubral courts—and won all three using *their* precedents against them. \\
\hline
Sister Oria & Custodian of the Lantern-Rites House & Keeps the old ways alive on the border; her threshold is a sieve for ill intent—Queen's enemies who cross it forget why they came, remembering only sudden peace. & Has never drawn a blade; keeps a bowl of Vhasian tears by her door, collected from those who wept at her threshold. \\
\hline
Earl Morcar of the North & Leader of the northern earls & Sees the Queen as a placeholder until a 'true' blooded king returns; his sword belonged to a Utaran-overthrown monarch—he draws it when charters feel too soft. & Carries a shard of Lence's cathedral glass in his pommel; whispers to it before every moot. \\
\hline
\end{longtable}

\subsection*{Names of Viterra}
\label{sec:viterra-names}

Inspired by Anglo-Saxon and early Norman England: strong consonants, soft vowels, surnames that recall a hedge, a dike, or a parish stained with history. Many nobles still bear Utaranized names from the occupation, but the common folk use the old tongue—the tongue spoken before Lence fell.

\begin{longtable}{|p{0.24\textwidth}|p{0.24\textwidth}|p{0.22\textwidth}|p{0.2\textwidth}|}
\hline
\textbf{First Names (Male)} & \textbf{First Names (Female)} & \textbf{Surnames} & \textbf{Epithets} \\
\hline
Baldwyn & Veralyn & Everblood & \textit{the Law-Giver} \\
\hline
Cuthbert & Aldric & Fenwood & \textit{the Unbuckled} \\
\hline
Roran & Halric & of the Scar & \textit{Bell-Sworn} \\
\hline
Morcar & Oria & de la Marche & \textit{the Hedge-Wise} \\
\hline
Leofric & Aelfwyn & l'Argent & \textit{the Silent} \\
\hline
Ghiro & Celandine & Riven & \textit{Ash-Banner} \\
\hline
\end{longtable}

\paragraph*{(Reeve/Guild/Knight/Legal Advocate)} Fen reeve with tally-rod speaking for dike guilds; river-carter syndic moving grain faster than rumor; justices of the peace settling disputes with careful pen—and a wary eye for Vhasian poison in the ink; master advocates who duel with words and swords, knowing the next word might be a blade.

\section*{Spades --- Places}
\label{sec:viterra-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Fen Causeway} --- Stile with toll-rod and eel carts queueing; each reeve claims right to collect under a different charter. Dueling advocates argue cases on makeshift platforms slick with river-mist and old blood. `[LAW]` `[TRADE]`
  \hfill \textit{The causeway sinks an inch each year. The lawyers argue over whose fault it is—and whether the sinking is a sign from Lence's drowned kings.}
\item \textbf{Hedgerow Green} --- In the Dales; bows strung under apple trees where justices of the peace hold court beneath makeshift arbors. Court duels permitted for matters of honor—and old scores. `[LAW]` `[COMBAT]`
  \hfill \textit{The apple tree is called ``the Hanging Judge.'' Its fruit tastes of iron—not from the soil, but from the Vhasian claimant hanged from its branches for poaching Queen's deer.}
\item \textbf{Beacon Hill} --- Above the Highlands; fire-grate still warm from last winter's levy call, signal tower with semaphore arms for legal dispatches and military alerts—each flash a warning or a lie. `[AUTHORITY]` `[TRAVEL]`
  \hfill \textit{The hill has three beacons. One for Ubral raids—lit when the cairns glow red. One for Vhasian incursions—lit when the Belworth runs thick with smugglers. One for paperwork—lit when the Queen's justiciars are lying.}
\item \textbf{Belworth Stairs} --- Ferry ledgers gone damp with river-mist; three manors claim toll rights, writs in three different hands—each a different shade of betrayal. Legal challenges fought in boats while waiting for the tide—or the assassin. `[LAW]` `[TRAVEL]`
  \hfill \textit{The stairs were repaired last year. The Vhasian side uses a different measurement—and a different kind of truth.}
\item \textbf{Bloom Quarry} --- Old iron-bloom turned drill yard; hammer echoes carry like war drums, forge-reeve holds tally of crown contracts—each one a hard-won truce. Military contracts awarded through legal bidding wars where the losers often 'find better jobs' face-down in the dirt. `[CRAFT]` `[AUTHORITY]`
  \hfill \textit{The forge-reeve has never lost a contract. His opponents have a habit of vanishing the night before the bid—leaving only their seals behind, still warm.}
\item \textbf{Parish Maze} --- Stone maze where three maps disagree; hedge-priest's vote decides who pays tithe and who holds freehold—because God, unlike men, does not lie in ledgers. Legal disputes settled by navigating the maze correctly—or by accepting the priest's blessing and walking away. `[LAW]` `[RIDDLE]`
  \hfill \textit{The maze changes at night. The hedge-priest says that's God will—and the lawyers say that's when the Vhasian smugglers move their goods.}
\item \textbf{Fairport Works} --- Tideworks at the Dolmis mouth; customs writs in three languages flutter like wounded birds, inspectors with magnifying glasses examine every manifest—for truth, or for the smuggler's tell. Maritime law enforcement includes boarding actions where the Queen's men often find Vhasian coins stamped with Viterran seals. `[TRADE]` `[AUTHORITY]`
  \hfill \textit{One language is Utaran—the language of the occupiers. One is Linn—the language of the traders. One is for the smugglers—written in salt and silence.}
\item \textbf{Law Quarter} --- Valora archives, oath-rooms, Hall of Dawning where precedent is birthed in ink and sealed with wax—each seal a promise, or a threat. Adjacent training grounds for legal advocates who practice their strikes on pellicles wrapped in old charters. `[LAW]` `[KNOWLEDGE]`
  \hfill \textit{The archives have a locked wing. The key is a legal argument no one has won—and no one dares to lose, for fear of what the winning side might do next.}
\item \textbf{Counting Fields} --- Tarlington beside muster ground; wagons load, ledgers mis-sum—not from error, but from the Vhasian poison slipped into the ink by desperate clerks. Auditors with quill-pens sharp as daggers and suspicious eyes; military mustering follows strict legal procedures with armed oversight—because numbers lie, and men kill for them. `[AUTHORITY]` `[LABOR]`
  \hfill \textit{The auditors carry swords. They say it's for counting cattle. No one asks further—but they watch the hands that hold the pens.}
\item[J] \textbf{Dike Crown} --- Ring of turf and timber; brotherhood bells warn of flood—and spies who move like water through the fen, carrying tidings of Vhasian raids or Queen's doubts. Legal sanctuary for those fleeing court judgments—if they can outrun the reeve's wrath. `[COMMUNITY]` `[LAW]`
  \hfill \textit{The sanctuary lasts for one tide. After that, the reeves can come—if they can find you before the Vhasian raiders do.}
\item[Q] \textbf{Progress Camp} --- Queen's canvas palisade, banners stiff, whispering old courtiers plot advancement through legal maneuver and sanctioned duels—each bout a test of whether the throne is steel or silk. `[AUTHORITY]` `[SOCIAL]`
  \hfill \textit{The camp moves every week. The lawsuits follow at a measured pace—each one a step closer to the Queen proving she is more than Fenwood's shadow.}
\item[K] \textbf{Tilt-Yard} --- Hall of Dawning at first light; Dawn-knights drill clean, loyalties in contracts signed with careful pen—each signature a vow, or a lie. Legal challenges often end in trial by combat—where the Queen's Peace is tested not in words, but in bruises. `[COMBAT]` `[LAW]`
  \hfill \textit{The sand is raked between duels. Sometimes the rake turns up old blood—or old oaths sworn on dying breath.}
\item[A] \textbf{Highway Stone} --- Queen's milestone above the Dolmis road; three sheriffs post seals over hers, each claiming different jurisdiction—and each seal a different kind of knife. Territory disputes resolved through armed arbitration—where the losers often don't walk away. `[AUTHORITY]` `[TRAVEL]`
  \hfill \textit{The stone has been broken and reglued seven times. Each crack holds a forgotten compromise—and a whispered oath from someone who thought they could cheat the tide.}
\end{enumerate}

\section*{Hearts --- People \& Factions}
\label{sec:viterra-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Fen Reeve} --- Tally-rod; speaks for dike guilds, remembers every flood-year's debts—and the charters that were bought with Vhasian gold or Queen's mercy. Authorized to call local levies for legal disputes—and to levy a blood-price if the papers don't add up. `[AUTHORITY]` `[COMMUNITY]`
  \hfill \textit{His tally-rod has notches from three floods. The next notch will be from a Vhasian raid he didn't report—to see if the Queen still trusts her reeves.}
\item \textbf{River Syndic} --- Moves grain faster than rumor; knows which writs matter—and which customs officials can be persuaded with a smile, a secret, or a slip of Vhasian silver. Maintains armed merchant guards who watch for Queen's men—or Vhasian knives in the grain. `[TRADE]` `[STEALTH]`
  \hfill \textit{His ledger has a watermark—a crown on one side, a sunburst on the other. He shows the right side to the right audience—and keeps the other for his own reckoning.}
\item \textbf{Parish Surveyor} --- Three maps, one opinion—but the opinion changes with the wind, and the tide, and the weight of a Vhasian coin in the pouch. Hedge-walks settle more disputes than courts—because the land remembers what the ledgers forget. Expert witness in property disputes—and in the quiet moments when a charter feels like a trap. `[KNOWLEDGE]` `[LAW]`
  \hfill \textit{His chain is exactly one rod long—but the rod changes length depending on who is paying—and whether they're paying in Vhasian silver or Queen's trust.}
\item \textbf{Dawn Quartermaster} --- Ledgers sharper than tongue; logistics before lances, counts every farthing spent in the Queen's name—each farthing a vote of confidence, or a bribe. Legal authority over military supplies—and over the Queen's own hesitation when the ledgers don't balance. `[AUTHORITY]` `[LABOR]`
  \hfill \textit{He once tracked a missing farthing to a Vhasian counting house. The farthing was returned—with interest, and a note in Vhasian script: ``Next time, we take the whole sack.''}
\item \textbf{Levy Serjeant} --- Dales bow calm, cider breath, memory keen for every loophole in the mustering laws—each loophole a place where Vhasian blood might soak into the soil. Commands armed civilians in legal enforcement actions—and in the quiet moments when a writ feels like a noose. `[COMBAT]` `[LAW]`
  \hfill \textit{His bowstring is made from a cancelled writ—one that promised peace, and delivered only war. He says it shoots truer now—because it remembers the cost of lies.}
\item \textbf{Cleric-Pair} --- Light preacher vs. Everflame canon lawyer; debates decide tithe law—not for the spirit, not for the letter, but for the quiet soul caught in the middle. Religious law carries weight in secular courts—because even kings fear a god who keeps score. `[SOCIAL]` `[LAW]`
  \hfill \textit{They share a room at the inn. They have not spoken outside court in seven years—but they still pass the salt with care, remembering the last time they did.}
\item \textbf{Fairport Shipwright} --- Dolmis cousins; skiff carries more secrets than cargo—and knows which manifests to forge for the Queen, and which to honor for the Vhasian ghost. Expert in maritime law and naval combat—where the truth is often in the hold, and the lie in the manifest. `[CRAFT]` `[STEALTH]`
  \hfill \textit{Her skiff has no name. The customs clerks have given it three—and each one is a different kind of truth.}
\item \textbf{Fenwood Comptroller} --- Conjures wagons with a signature—each signature a promise, or a plea. Silence with a seal—each seal a different kind of truth. Counts the Queen's coin—and tracks every debt owed to the crown—each debt a different kind of wound. Legal authority over taxation—and over the Queen's own guilt when the books don't balance. `[AUTHORITY]` `[TRADE]`
  \hfill \textit{His quill is a dagger. He has used it as both—to write, and to remind himself what happens when the ink runs dry.}
\item \textbf{Queen's Justiciar} --- Law on the road, polite as a noose—but her grip is tight, and her eyes scan the tree line for Vhasian shadows—or the Queen's own doubt. Bearing three kings' judgments and a purse full of pardons for the right price. Authorized to conduct trial by combat—where the law is not in the words, but in the weight of the blade. `[LAW]` `[AUTHORITY]`
  \hfill \textit{She has never lost a legal duel. Her opponents have a tendency to fall down stairs—and then lie very still, wondering if they've wasted the Queen's time.}
\item[J] \textbf{Border Routier} --- Reads ledgers as ambushes—and writes them as tribute. Serves two crowns, collects from both, knows which border posts take bribes in what coin—and which take Vhasian silver for Queen's blood. Expert in frontier law enforcement—where the ledger is often the first thing to burn. `[STEALTH]` `[TRADE]`
  \hfill \textit{His loyalty is to the highest bidder. The Queen and the Vhasian claimant are currently in a bidding war—and the currency is lives.}
\item[Q] \textbf{Warrior Queen} --- Crown on helm, sword at side—she hasn't drawn it in months, but the weight is familiar. Patient sums behind her gaze; not strategy, but the quiet calculation of how much blood it will take to keep the peace. Final arbiter in legal duels—where she wins not by being right, but by being willing to go further. `[AUTHORITY]` `[LAW]`
  \hfill \textit{Her crown has no jewels—only seals. Each seal is a conquered claim—and each one is still warm from the hand that took it.}
\item[K] \textbf{Crown Council} --- Fenwood dukes, guild envoys, abbots; weighing grain against glory—and wondering if the Queen remembers why she took the throne. Charters against coin, precedent against practicality—and the shadow of Baldwyn Brightstide against them all. Legal battles fought through proxy advocates—because sometimes, the truest voice is the one that doesn't speak. `[POLITICS]` `[LAW]`
  \hfill \textit{The council has a standing rule: no weapons within the chamber. The advocates have found seventeen loopholes—and eighteen reasons to break it.}
\item[A] \textbf{Master Advocate Halric} --- Silver-tongued legal warrior who settles disputes through verbal duels that can end in physical combat—because sometimes, the only true law is the one written in blood. Keeper of legal combat traditions—and of the quiet hope that words might yet be enough. `[LAW]` `[COMBAT]`
  \hfill \textit{He has argued cases in Vhasian, Thepyrgos, and Ubral courts—and won all three using their own precedents against them—and their own blades when the words ran out.}
\end{enumerate}

\section*{Clubs --- Complications/Threats}
\label{sec:viterra-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Dike Breach} --- Black-rain; the wrong reeve claims credit while wagons drown—but the water tells another story, carrying not just peat, but the ghost of Lence's last stand downstream. Legal liability may require armed enforcement—and a silent prayer to the drowned kings. `[DISASTER]` `[LAW]`
  \hfill \textit{The water is brown with peat and lies. It carries old deeds downstream—and the screams of those who signed them.}
\item \textbf{Feast Clash} --- Light vs. Everflame processions collide—but it's not an accident. Shutters slam, writs fly, justices of the peace called to settle which holy day takes precedence—and which god gets the first blood. Honor disputes may require legal duels—and a quiet moment at the shrine. `[SOCIAL]` `[LAW]`
  \hfill \textit{The processions take different routes now. The clash was not an accident—it was an invitation.}
\item \textbf{Quiet Tolls} --- Sprout on the Queen's Highway—but it's not just sprouts. Escorts sniff traps laid not just by smugglers, but by the Queen's own doubts. Sheriffs argue rights. Customs officials demand duties no one knew existed—because the Vhasian routiers wrote them in blood last autumn. Territorial disputes resolved through armed arbitration—where the losers often don't get to argue. `[AUTHORITY]` `[TRAVEL]`
  \hfill \textit{The toll posts have fresh paint. The paint is from the keep of a Vhasian claimant—and it's still wet.}
\item \textbf{Audit Freeze} --- Counting-house freezes cargo—but it's not the cargo that's cold. It's the ink—slipped with Vhasian poison by desperate clerks hoping to buy time. Queen's seal vs. old writs, auditors with suspicious eyes and quill-pens sharp as daggers. Legal challenges may escalate to physical confrontation—and a quick check for poison in the well. `[AUTHORITY]` `[DEBT]`
  \hfill \textit{The auditors work by candlelight. The shadows on the walls look like gallows—and smell like Vhasian garlic.}
\item \textbf{Border Snarl} --- Three courts arrest the same hay-cart—but it's not about the hay. It's about whose charter was signed with cleaner hands—and whose soul is less stained. Whose hay is it? Which lord's charter takes precedence? Bribes flow like wine—and Vhasian silver. Frontier justice often involves armed standoffs—and a moment of silence for the dead. `[LAW]` `[COMBAT]`
  \hfill \textit{The hay is from a field that was Vhasian last spring. Now it's Viterran. The field hasn't moved—but the souls working it have.}
\item \textbf{Isle Refusal} --- Theona withholds levy—but it's not about the money. The quays choke on rumor not just of greed, but of the old curses the Three Greens mutter when Vritann boots touch their soil. Ships idle while lawyers debate maritime law and ancient treaties—and whether the Queen's peace is worth the salt in their wounds. Naval blockades become legal enforcement actions—where the truth is often in the bilge. `[POLITICS]` `[LAW]`
  \hfill \textit{Theona's envoy carries a hawthorn branch—not of peace, but of warning. The Queen's justiciars will not touch it—and will wonder why it doesn't burn.}
\item \textbf{Delta Spat} --- Fairport vs. Marcott customs—but it's not about the manifests. Barges stack three deep, merchants pick sides—not just for profit, but for the old scores. Inspectors with magnifying glasses examine every manifest—for truth, or for the Vhasian tell. Trade disputes fought with merchant guards—who often find Queen's men in the hold, and Vhasian coins in the manifest. `[TRADE]` `[LAW]`
  \hfill \textit{The magnifying glasses reveal truth—and also start fires. Both are useful—but only one keeps the peace.}
\item \textbf{Routier Arrears} --- A free-company flips unless someone pays—but it's not about the money. It's about the name on the ledger—and whether it's written in Vhasian silver, or Queen's trust. Your name on it—and the debt owed to men who know the price of blood—and the quiet terror of owing it to the wrong side. Legal debts enforced through hired swords—who often collect more than coin. `[DEBT]` `[COMBAT]`
  \hfill \textit{The routiers have a list. Your name is on it—or will be by sunset—and the ink might be Vhasian, or might be Queen's, and either way, it's a debt written in blood.}
\item \textbf{Salt Pinch} --- Import spikes—but it's not just about the salt. Bakers barricade—not just for profit, but for the old scores. The Queen's promise tastes like thin gruel—not because of the salt, but because of the ash from the burned Vhasian customs post—and the quiet fear that the peace is already broken. Merchants hoard salt like gold—not for trade, but for the coming storm. Riots brew in back alleys—not just for hunger, but for the chance to settle the score. Economic warfare becomes physical conflict—where the first blow is often struck over a pinch of salt. `[WEATHER]` `[SOCIAL]`
  \hfill \textit{The salt is grey. It was cut with ashes from a burned Vhasian customs post—and it still smells like smoke.}
\item[J] \textbf{Dawn Recall} --- Escort pulled to a flood-struck parish—but it's not just about the flood. The writ argues with the river—and the river argues back, carrying not just water, but the Queen's doubts. Justices demand inquiry into who neglected the dikes—and who let the Vhasian raiders through. Military enforcement of legal judgments—where the enforcement often looks suspiciously like a raid. `[AUTHORITY]` `[TRAVEL]`
  \hfill \textit{The escorts are back in a week. They bring new writs—and fewer men—and more questions about who they're really serving.}
\item[Q] \textbf{Aberielist Intrigue} --- Royalist seals in wrong hands—but it's not just about the seals. It's about the whispers in the counting houses—and whether the Queen's patience has finally snapped. Conspirators plot in the quiet places, where the law feels thin—and the blood feels close. Treason charges carry death penalty in legal duels—where the truth is often in the silence between words. `[POLITICS]` `[LAW]`
  \hfill \textit{The seals are from a dead duke—and the duke's ghost is said to approve. Or to weep.}
\item[K] \textbf{Levy Call-Up} --- Dalesmen mustered—but it's not just about the wagons. It's about the soul behind the plow—and whether they still believe in the Queen's peace. Wagons seized ''for the realm,'' conscription writs signed with careful pen—each signature a different kind of promise. Families left to tend fields alone—and to wonder if the Queen still believes in them. Legal authority backed by armed force—where the force often looks like the Queen's own doubt made manifest. `[AUTHORITY]` `[COMBAT]`
  \hfill \textit{The conscription writs have a clause: ``void if Belworth freezes.'' The river is still warm—but the Queen's hand sometimes shakes when she signs it.}
\item[A] \textbf{Legal Duel Challenge} --- Master Advocate Halric issues a formal challenge—but it's not just parchment. It's a gauntlet thrown in the tilt-yard at first light, where the sand still holds the blood of last week's duel. Refuse and lose legal standing—accept and risk life and fortune in court-yard combat—where the winner takes all, and the loser learns the cost of doubting the Queen's peace. The ultimate legal weapon—and the first test of whether the throne is steel or silk. `[LAW]` `[COMBAT]`
  \hfill \textit{The challenge is written on both sides of the parchment—and there is no blank space to refuse. Or to run.}
\end{enumerate}

\section*{Diamonds --- Rewards/Leverage}
\label{sec:viterra-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Ferry Priority} --- Belworth crossing (once); a horn answers—but it's not just any horn. It's blown by a Vhasian defector who now serves the Queen, and the note is the same one that sounded the day Lence fell. Legal right of way backed by armed escort—and the quiet hope that the peace is real. `[TRAVEL]`
  \hfill \textit{The priority token is a writ signed by both banks. The ink on one side is newer—and the other side still smells of Vhasian smoke.}
\item \textbf{Dike Allotment} --- Brotherhood labor granted—but it's not just about the work. It's about whether three clerks can look past the old scores—and whether the bribes distributed among the right hands are clean, or still carry the ghost of Lence. Work performed under legal contract with military oversight—and the quiet watch of the Fenwood dukes. `[LABOR]`
  \hfill \textit{The allotment includes a clause: ``as tides permit.'' The reeve decides what the tides permit—and whether the Queen's trust is worth the risk.}
\item \textbf{Market License} --- Valora's square—but it's not just about the square. It's about whether a guild nod is enough—and whether the right seal on the right parchment opens doors, or just invites trouble. Legal protection from merchant guild enforcement—and the quiet irony of being protected by a Vhasian abbey's seal. `[TRADE]`
  \hfill \textit{The license is written on vellum from a Vhasian abbey. The irony is noted—and the ink is still wet.}
\item \textbf{Dawn Escort} --- Four lances at dawn—but it's not just about the lances. It's about whether loyalty waxes with pay—or wanes with the Queen's doubt. Their presence keeps lesser men from troubling you—and the Queen from wondering if she's still worthy. Legal authority to enforce your writs—and to remind the realm that the peace is real. `[AUTHORITY]`
  \hfill \textit{The lances are from the Fenwood armory—and the Fenwood dukes have their own plans—and their own doubts.}
\item \textbf{River Charter} --- Guaranteed haul on a grain artery—but it's not just about the haul. It's about whether the weather permits—and whether the duties paid to the right officials are clean, or still carry the Vhasian ghost. Legal right protected by river patrol—and the quiet hope that the peace is worth the salt. `[TRAVEL]`
  \hfill \textit{The charter is stamped with a seal that looks like a crown—or a sunburst. No one can agree—but everyone agrees the ink should be dry.}
\item \textbf{Map Correction} --- A hedge moves—but it's not just about the hedge. It's about whether the border shifts with the truth—or with the old scores. Lawsuits bloom—not just for profit, but for the chance to settle the score. A new survey can make or break a man's fortune overnight—and his soul in the process. Legal survey backed by armed surveyors—who carry chains, swords, and the quiet hope that the truth is worth the blood. `[LAW]` 
  \hfill \textit{The surveyors carry chains and swords. The swords are for when the chains disagree—and for when the truth is worth the blood.}
\item \textbf{Customs Seal} --- Fairport for Dolmis cargo—but it's not just about the seal. It's about whether it's honored today—and disputed tomorrow, or whether it's a quiet truce between old scores. Legal immunity from customs enforcement—and the quiet hope that the peace is worth the salt in the wound. `[TRADE]`
  \hfill \textit{The seal is a wax crab. It pinches when pressed—and it remembers the hand that shaped it.}
\item \textbf{Thing Ruling} --- Sharp local bite—but it's not just about the bite. It's about whether appeals run swifter than oaths—and whether justice is swift, or just quiet. Court decisions enforceable by local levy—and the quiet hope that the ruling is worth the blood in the soil. `[LAW]` 
  \hfill \textit{The ruling is carved on an oar. Break the oar, and the ruling breaks with it—and the soil remembers.}
\item \textbf{Salt Allotment} --- Winter depot—but it's not just about the depot. It's about whether the Queen's writ opens doors—and whether the old hands still count sacks, or if they've learned to trust the Queen's promise. The right connections ensure the best salt reaches the right tables—and the quiet hope that the peace is worth the blood in the wound. `[TRADE]` 
  \hfill \textit{The salt is from a pan worked by Ubral exiles—and they still count in clan measure—and in the quiet hope that the peace is worth the blood.}
\item[J] \textbf{Wardship} --- A minor fen-keep—but it's not just about the keep. It's about whether the men-at-arms are ''for now,'' or if they're ready to stand. Whether the tutor has plans—and whether the right guardian can be found for the right price, or if they're still waiting for the Queen to decide. Legal guardianship with military authority—and the quiet hope that the child is worth the trust. `[AUTHORITY]` 
  \hfill \textit{The ward is a child. The child has a seal ring—and the ring is older than the kingdom—and older than the doubt.}
\item[Q] \textbf{Queen's Audience} --- Private—but it's not just about the audience. It's about whether her questions strike like thrusts—and whether her favors can elevate a man from nothing to something, or if they're just a quiet test of whether he's worth the risk. Royal decree carries immediate enforcement power—and the quiet hope that the peace is real. `[AUTHORITY]` 
  \hfill \textit{The audience lasts one hour. The paperwork lasts a generation—and the Queen's doubt lasts longer.}
\item[K] \textbf{Ducal Warrant} --- Fenwood to enforce customs—but it's not just about the warrant. It's about whether old power bows—and whether new power watches, or if they're still waiting for the Queen to decide if she's still theirs. The parchment carries weight in any court—and the quiet hope that the peace is worth the blood in the ink. `[AUTHORITY]` 
  \hfill \textit{The warrant is signed in blood—or ink that looks like blood. The difference matters at trial—and in the quiet moments when the Queen wonders if she's still Viterra's queen.}
\item[A] \textbf{Duel Sanction} --- Right to challenge in legal combat—but it's not just about the right. It's about whether the Queen's seal makes it official—and whether the Dawn-knights serve as seconds, or if they're still waiting for her to decide if she's worth the risk. The winner takes all legally—and the loser learns the cost of doubting the Queen's peace. Ultimate legal enforcement mechanism—and the first test of whether the throne is steel or silk. `[LAW]` `[COMBAT]`
  \hfill \textit{The sanction is a coin. Heads you duel, tails your opponent chooses the weapon—and the coin is still warm from the last time it was flipped in earnest.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:viterra-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified outcomes (writs/charters/priority) that change position rather than call for a roll.
\item If any A appears, echo \textbf{hedge-law} motif—boundaries blur like old wounds, charters contradict themselves like guilty consciences, precedent splinters like shaky oaths, and the careful web of legal authority shows its cracks—not as weakness, but as the living pulse of a realm still deciding what it means to be free. Legal warfare escalates to physical conflict—not as failure, but as the only language left when words run dry.
\end{itemize}

\subsection*{Additional Features (Cultural Mechanics)}
\label{sec:viterra-mechanics}

\begin{itemize}
\item \textbf{Hedge-Law Momentum:} When a valid \emph{writ, charter, or warrant} is presented in the proper parish—but only if the ink is dry and the seal is unbroken—step \textbf{Position} up one category on social or logistical actions in that locale for the scene. The first contradiction starts a \textbf{Border-Lace} [4] clock; advance each time competing papers are invoked—and each time the Queen's hand shakes when she signs them.
\item \textbf{Beacon \& Bell Chain:} If two adjacent \emph{beacons} or \emph{brotherhood bells} are under your influence—and if they're lit for the right reason, not just habit—convert one \textbf{Club} draw into a travel delay for \emph{others} while the party proceeds with improved \textbf{Effect} for movement and logistics. The signal draws scrutiny; the GM may bank 1 SB tied to local authority—and to the Queen's quiet doubt.
\item \textbf{River-Right Window:} At a ferry, tidework, or customs stair—if the water runs clear and the air smells of salt, not smoke—play any \textbf{Diamond} tied to crossings or seals to open a short \emph{window of lawful passage}: one convoy passes uninspected, or one cargo clears at priority. The window closes immediately if a storm, audit, or levy card is revealed—or if the Belworth runs red with old blood.
\item \textbf{Legal Duel Protocol:} When legal disputes cannot be resolved through argument—and when the words feel too thin for the weight of the claim—parties may challenge each other to a duel. The winner's position becomes legally binding. Requires proper sanction and seconds (see \textbf{Duel Sanction} Diamond)—and the quiet courage to believe that words might yet be enough.
\item \textbf{Vhasian Claims:} Any Viterran noble with land across the Belworth may invoke a \emph{Ducal Warrant} to attempt to collect rents or enforce judgments in Vhasia—but only if they're ready to face the ghost of Lence in the other's eyes. Doing so starts a \textbf{Border Snarl} clock, as the Vhasian claimants are unlikely to cooperate—and the Queen is unlikely to sleep.
\item \textbf{Theona Contest:} The Warrior Queen's authority over the Three Greens is nominal at best—but it's not just about the crown. It's about whether she's still worthy of it. Any attempt to enforce Viterran law in Theona triggers an \textbf{Isle Refusal} complication, and the local taboos may twist the law into knots—and the Queen's doubt into resolve.
\item \textbf{The Succession's Shadow:} The sudden death of King Baldwyn Brightstide—and the quiet whisper that it was Fenwood's doing—left deep scars. Any legal dispute that touches on inheritance, legitimacy, or royal authority is automatically one step more volatile. The GM may spend 1 SB to introduce a witness, a document, or a rumor related to the succession crisis—and to wonder if the Queen still believes in her own right to rule.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{Saint Elodie of the Unburdened Knot (Inaea / Oath of Mercy and Grace):} A legendary figure from the early days of Viterran independence—not a priestess, but a hedge-walker who knew that some debts are too heavy to carry. She wandered the disputed borderlands, binding feuds with red thread and forgiveness. Her most famous miracle was the \emph{Unknotting of the Fenwood Vengeance}—a blood feud that had run for nine generations. She took a strand of hair from each family's eldest, tied them into a single cord, and then \emph{cut the cord in half}, declaring: ''The past is severed. What you owe now is not to the dead, but to the next stranger who asks for bread.'' The families feasted together for the first time in centuries. Her shrines are wayside posts where travellers leave a knotted cord; the first person to untie a knot and offer it to a stranger in need is said to receive her blessing (+1 die to their next act of mercy or reconciliation). The Temple of Light has never canonised her—she was too folk, too female, too hedgerow—but the common folk remember. The Queen's justiciars respect her precedent: \emph{A severed debt is not a gift. It is a thread that now binds you to the living.}
    \item \textbf{Saint Michel of the Gilded Scale} — Mykkiel. *The law that does not bend; the judge who cannot be bought.*
    \item \textbf{Saint Isabeau of the Lilies} — Ikasha. *The dream of setting aside the crown; the mercy of forgetting duty.*
    \item \textbf{Saint Evroul of the Cairn} — Mor'iraath. *The oath that outlasts the oath‑taker; the debt paid in blood and stone.*
    \item \textbf{The Hedge King} — The Traveler. *The road that belongs to no lord; the path that finds its own way.*
    \item \textbf{The Brightstide Martyr} (folk cult) — A local veneration of King Baldwyn Brightstide, who died before he could rule. His followers believe his ghost still rides the borders—and that the Queen rules in his name—not her own. And they wonder, in the quiet hours, if she believes it herself.
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross‑Expansion Hooks)}
\label{sec:viterra-lore}
\begin{itemize}
  \item \textbf{The Hedge-Seal (The Dike Covenant):} Covenant judges in the eastern parishes adjudicate disputes over river tolls and inheritance—but not just with scales and tables. They carry the weight of the Belworth in their bones, and know that some truths are too old for parchment. They carry no weapons, only a set of iron scales and a copy of the \emph{Hedge-Law Tables}—each page stained with the water of a different flood. A ruling from the Order is binding in both Viterra and the Vhasian marches—but any party who lies before them forfeits their right to appeal—and their peace of mind. The Warrior Queen has granted them a charter; the Fenwood dukes consider them spies—and wonder if they're still loyal to the old ways.
  \item \textbf{Saint Cuthbert of the Fen (The Traveler):} A legendary reeve who walked the flooded dikes during the Great Storm—and lit every beacon by hand, not just duty, but hope. His shrines are stiles and bridge posts; travellers who touch the iron ring on his shrine and speak a destination find the road holds firm—and the quiet hope that the peace is real. The Temple of Light has absorbed his cult, but the fen folk still leave offerings of eels at his cairns—and wonder if the Queen still believes in her own mercy.
  \item \textbf{The Cog-Cults of the Northern Hills (The Clockwork Monad):} In the high parishes near the Ubral border, secret fraternities of millers and surveyors believe that law is a machine that can be perfected—but they wonder if the machine is worth the soul. They meet in abandoned watermills, measuring boundaries with brass chains and reciting the \emph{Canons of Proportion}. The Everflame calls them heretics; the Queen's justiciars ignore them—because they're still waiting for her to decide if she's worth the trust.
  \item \textbf{The Salt Sworn of the Eastern Coast (The Pale Shepherd):} Along the Dolmis shore, a syncretic cult tends to the drowned—but they wonder if the sea is worth the tears. They build cairns of salt stone and leave no offerings—only the names of the lost, carved into driftwood. The Temple of Light has tried to suppress them three times—each time, the mission ships have been found adrift with no crew and a single name carved into the mast: \emph{Oblivion}. And they wonder, in the quiet hours, if the Queen still believes in her own oblivion.
  \item \textbf{The Ninth:} In any Viterran court, the ninth citation of precedent voids all previous citations—but it's not just about the law. It's about the quiet doubt that maybe the peace isn't worth the blood. The wise stop at eight and settle out of court. Those who cite the ninth are said to hear a faint laughter from the marble walls—and wonder if it's the Queen's laugh, or the ghost of her brother's.
\end{itemize}

\subsection*{Viterran Superstitions (burn for +1 die once)}
\label{sec:viterra-superstitions}
\begin{itemize}
  \item \textbf{Bread to the Hedge:} Crumble a crust at a parish boundary stone—but it's not just about the stone. It's about whether the hedge still holds—and if the Queen still believes in her own hedge. Ignore the first \emph{Border Snarl} penalty—and wonder if the peace is real. `[SUPERSTITION]`
  \item \textbf{Knot the Writ:} Tie a single knot in a charter's ribbon—but it's not just about the ribbon. It's about whether the promise is worth the ink—and if the Queen still believes in her own word. Cancel \emph{Audit Freeze} or take --1 die on your next legal argument (your choice)—and wonder if the doubt is worth the risk. `[SUPERSTITION]`
  \item \textbf{Name to the River:} Whisper a dead claimant's name into the Belworth—but it's not just about the name. It's about whether the river still remembers—and if the Queen still believes in her own right to rule. Gain +1 Effect when arguing against their heir's claim this scene, but mark \emph{Obligation} to the Queen's Justiciar—and wonder if the obligation is worth the blood. `[SUPERSTITION]`
  \item \textbf{The Brightstide Candle:} Light a candle in memory of King Baldwyn before a difficult legal proceeding—but it's not just about the candle. It's about whether the ghost still walks—and if the Queen still believes in her own right to the throne. On a roll that includes a 1, you may reroll that die once—but if you still fail, the ghost of the dead king is said to watch from the shadows—and wonder if she's still worth the shadow. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:vitarra-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{crown officials or rival advocates—but it's not just about the eyes. It's about the quiet doubt in their gaze, and whether they still believe in the Queen's peace.}
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{writ‑runners or levy guards—but it's not just about the sound. It's about the silence that follows, and whether it's filled with the ghost of Lence.}
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{trackers or customs inspectors—but it's not just about the trace. It's about whether the ink is dry, and if the Queen still believes in her own word.}
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable court date—but it's not just about the time. It's about whether the peace is worth the wait.}
\item \textbf{Supply Strain:} Mark +1 segment on a relevant \textbf{resource clock}. 
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Queen's Justiciar or fen reeve} becomes aware and begins responding—but it's not just about the alarm. It's about whether the Queen still believes in her own right to call them.}
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{legal challenge or river closure—but it's not just about the ground. It's about whether the hedge still holds, and if the Queen still believes in her own hedge.}
\item \textbf{Foe Appears:} A \textbf{rival claimant or border routier} arrives on scene—but it's not just about the foe. It's about whether they still believe in the Queen's peace—and if she still believes in herself.}
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.}
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.}
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{Dawn‑knights, levy forces, or Vhasian routiers} arrive—but it's not just about the reinforcements. It's about whether the Queen still believes in her own right to command them.}
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.}
\item \textbf{Major Twist:} The situation fundamentally changes — \textbf{writ voided/duel declared/levy called—but it's not just about the twist. It's about whether the peace is worth the blood.}
\item \textbf{Rail Tick:} Advance a relevant campaign/front clock by 1 segment.}
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.}
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect—but it's not just about the trap. It's about whether the peace is worth the blood.}
\item \textbf{Authority Arrival:} \textbf{Warrior Queen, Crown Council, or Master Advocate Halric} intervenes—but it's not just about the authority. It's about whether the Queen still believes in her own right to rule.}
\item \textbf{Scene Shift:} The environment changes dramatically — \textbf{dike breaks/river floods/court erupts—but it's not just about the shift. It's about whether the peace is worth the blood.}
\item \textbf{Patron Omen:} Divine/arcane forces take notice — \textbf{omen appears/blessing lost/curse manifests—but it's not just about the omen. It's about whether the peace is worth the blood.}
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives—but it's not just about the pivot. It's about whether the peace is worth the blood.}
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Viterra (Hedge Law):} Boundary stones shift like old scars, charters contradict themselves like guilty consciences, legal precedents split mid-citation like shaky oaths.}
\item \textbf{Viterra (Cross‑Border Claims):} Vhasian landlords appear with older deeds—and the quiet hope that the peace is worth the blood. Theonan taboos void writs—and the Queen wonders if she's still worth the trust. Ubral cairns declare neutrality with axes—and the Queen wonders if she's still worth the fear.}
\item \textbf{Viterra (Legal Warfare):} Court decisions reverse without warning—and the Queen wonders if she's still worth the trust. Witnesses recant—and the Queen wonders if she's still worth the doubt. Duel seconds change sides—and the Queen wonders if she's still worth the risk.}
\item \textbf{Viterra (Succession Politics):} A document surfaces bearing King Baldwyn's seal—and the Queen wonders if she's still worth the shadow. A northern earl declares the Queen's writ invalid in his lands—and the Queen wonders if she's still worth the trust. A baron offers to switch sides—for a price—and the Queen wonders if she's still worth the cost.}
\end{itemize}

\subsection*{Pursuit Ladder (Near / Mid / Far)}
\label{sec:viterra-pursuit}
\begin{itemize}
\item \textbf{Advance/Withdraw (Wits + Survival):} On a hit, shift one band. On a strong hit, also impose \emph{Thorn Hedge}.}
\item \textbf{Cut the Charter (Presence + Sway):} On a hit, freeze range for one exchange; if a \textbf{Ducal Warrant} is in your possession, +1 die—but only if you still believe in the Queen's peace.}
\item \textbf{Weather the Assize (Resolve + Lore):} On a hit in \emph{Audit Freeze}, you pick whose ledgers are examined; on a miss, the GM does—and wonders if the Queen still believes in her own right to doubt.}
\end{itemize}

\subsection*{Legal Combat System (Summary)}
\label{sec:viterra-legal-combat}
When legal disputes cannot be resolved through negotiation—and when the words feel too thin for the weight of the claim—parties may engage in formal legal duels. These combine verbal sparring with physical combat—where the truest law is often written in blood.

\textbf{Phase 1: Pleadings}
\begin{itemize}
\item Each advocate presents their case (Wits + Sway vs. DV 3)—not just with words, but with the weight of their belief in the Queen's peace.}
\item Success grants +1 die to the next phase—and the quiet hope that the words were enough.}
\item Failure creates 1 SB that the opponent can spend—and the quiet doubt that maybe the sword was needed all along.}
\end{itemize}

\textbf{Phase 2: Cross-Examination}
\begin{itemize}
\item Attack opponent's arguments (Wits + Insight vs. DV 3)—not just with logic, but with the weight of their belief in the Queen's peace.}
\item Success inflicts Fatigue 1 on the opponent—and the quiet hope that the doubt was worth the cost.}
\item Critical success forces opponent to answer truthfully—and the quiet hope that the truth is worth the blood.}
\end{itemize}

\textbf{Phase 3: Trial by Combat (if unresolved)}
\begin{itemize}
\item Physical combat with legal stakes (Body + Melee or relevant combat skill)—not just for victory, but for the quiet hope that the peace is worth the blood.}
\item Winner's position becomes legally binding—and the Queen wonders if she's still worth the trust.}
\item Loser suffers legal consequences (fine, loss of claim, imprisonment) plus physical harm—and the Queen wonders if she's still worth the doubt.}
\end{itemize}

\subsection*{Boss Seeds (three‑phase foes)}
\label{sec:viterra-bosses}
\begin{itemize}
\item \textbf{The Violet Clerk (Master of Counter‑Seals)} — \emph{Phases}: Produce Earlier Patent $\rightarrow$ Split the Writ $\rightarrow$ Seal Fast. \emph{Cracks}: a \textbf{Queen's Audience} overrules his seal—and the Queen wonders if she's still worth the trust; a \textbf{Duel Sanction} forces personal combat—and the Queen wonders if she's still worth the risk; a \textbf{Map Correction} voids his boundary claim—and the Queen wonders if she's still worth the doubt.}
\item \textbf{The Thorn Captain (Queen's Quiet Hand)} — \emph{Phases}: Queenspeace Zone $\rightarrow$ Duel Warrant $\rightarrow$ Swap the Field. \emph{Cracks}: a \textbf{Map Correction} reveals a loophole—and the Queen wonders if she's still worth the doubt; a \textbf{Dike Allotment} lets you claim the field as ``flood zone''—and the Queen wonders if she's still worth the trust; a \textbf{Border Routier} testifies to the Queen's favor—and the Queen wonders if she's still worth the risk.}
\item \textbf{The Saffron Abbess (Keeper of Bell‑Law)} — \emph{Phases}: Wrong Hour $\rightarrow$ Bless or Damn $\rightarrow$ Mixed Tribunal. \emph{Cracks}: a \textbf{Queen's Justiciar} invokes crown supremacy—and the Queen wonders if she's still worth the trust; a \textbf{Border Routier} bribes her bell‑ringer—and the Queen wonders if she's still worth the risk; a \textbf{Thing Ruling} from Theona forces her to recuse—and the Queen wonders if she's still worth the doubt.}
\item \textbf{The Brightstide Claimant (Pretender)} — \emph{Phases}: Produce the Lost Charter $\rightarrow$ Rally the Northern Earls $\rightarrow$ March on Valora. \emph{Cracks}: find proof that the charter is a forgery—and the Queen wonders if she's still worth the trust; convince Earl Morcar to stand with the Queen—and the Queen wonders if she's still worth the risk; survive a legal duel with the claimant's champion—and the Queen wonders if she's still worth the doubt.}
\end{itemize}

\subsection*{Legendary Sites \& Trials}
\label{sec:viterra-trials}
\begin{itemize}
\item \textbf{Belworth Stairs Ordeal} --- Cross the river without a charter—but it's not just about the crossing. It's about whether the Queen still believes in her own right to demand it. On success, you prove that possession is nine‑tenths of the law—and the quiet hope that the peace is real. On failure, the river takes your name—and your cargo—and the Queen wonders if she's still worth the doubt. (Wits + Survival vs. DV 5.)}
\item \textbf{Hall of Dawning Vigil} --- Spend a night in the tilt-yard—but it's not just about the night. It's about whether the Queen still believes in her own right to doubt. At dawn, you recall a forgotten legal precedent—and the quiet hope that the peace is real. Gain a one‑use \emph{Precedent Edge} (+1 die in any court). The ghost of a dead duelist may challenge you to a legal dream‑duel—and the Queen wonders if she's still worth the shadow. (Spirit + Lore vs. DV 4.)}
\item \textbf{Hedge‑Moot at Dusk} --- Settle a boundary dispute by walking the hedge—but it's not just about the walk. It's about whether the hedge still holds, and if the Queen still believes in her own hedge. Each successful navigation test (Wits + Survival) clears 1 segment from a \textbf{Border Snarl} clock—and the quiet hope that the peace is real. A miss plants the thorn of a future feud (advance \textbf{Border Snarl} by 1)—and the Queen wonders if she's still worth the doubt.}
\end{itemize}

% ====== VITERRA: PLUG‑IN EXTENSIONS ======
\subsection*{Extensions (Plug-in)}
\label{sec:viterra-extensions}

\begin{itemize}
  \item \textbf{Legal Dial (Crown $\leftrightarrow$ Local Custom):} Track the party's stance on authority—and on whether they still believe in the Queen's peace.
  \begin{itemize}
    \item \emph{Crown High:} +1 Position when invoking royal writs, charters, or the Queen's name—but only if they still believe in her right to rule; first appeal to local custom starts one step worse—and the Queen wonders if she's still worth the trust.}
    \item \emph{Local High:} +1 Effect on hedge‑law, parish rights, or ancient precedent—but only if they still believe in the old ways; first royal petition starts one step worse—and the Queen wonders if she's still worth the doubt.}
    \item Center the dial by hosting a mixed moot or surrendering a contested seal to a neutral third party—and wonder if the peace is worth the blood.}
  \end{itemize}

  \item \textbf{Courtroom Mini‑Game (fast trial):}
  \begin{itemize}
    \item Lay three stones: \textbf{Precedent}, \textbf{Equity}, \textbf{Force}.  
    \item Argue a point (roll Lore+Diplomacy vs. DV 3)—not just with words, but with the weight of their belief in the Queen's peace. On a hit, move one stone to your side. On a miss, the opposing advocate gains \emph{Citation Edge}.  
    \item Win two stones to secure a temporary \textbf{Court Ruling} (treat as a Diamond for one scene).  
    \item \emph{Halric's Interjection (2 SB):} flip \textbf{Precedent}, but start \textbf{Whispered Precedent} [2] with the losing side—and wonder if the Queen still believes in her own right to doubt.}
  \end{itemize}

  \item \textbf{Cross‑Border Tensions (fast pressure):}
  \begin{itemize}
    \item Two opposed clocks: \emph{Queen's Peace} vs \emph{Border Fire} (size by highest rank in the seed).  
    \item Each leg, choose \emph{Viterran Road} (safe, slow), \emph{Vhasian Claim} (fast, risky), or \textbf{Ubral Pass} (hidden, costly).  
    \item Advantage: Vhasian Claim$\rightarrow$Viterran Road (shortcuts), Ubral Pass$\rightarrow$Vhasian Claim (concealment), Viterran Road$\rightarrow$Ubral Pass (supplies).  
    \item Spend a \textbf{Ducal Warrant} to shift one segment from \emph{Border Fire} to \emph{Queen's Peace}—and wonder if the Queen still believes in her own right to rule.}
  \end{itemize}

  \item \textbf{Start‑of‑Scene Rumors (1d6):}
  \begin{enumerate}
    \item \emph{Wrong Wax:} a rival's seals softened; +1d to forgeries this scene—and the Queen wonders if she's still worth the trust.}
    \item \emph{Empty Beacon:} tower unmanned; travel faster, scrutiny harsher—and the Queen wonders if she's still worth the doubt.}
    \item \emph{Dawn‑Lane Cleared:} first convoy action gains +1 Position—and the Queen wonders if she's still worth the trust.}
    \item \emph{Clerk's Wedding:} filings delayed; first petition ignores \emph{Audit Freeze}—and the Queen wonders if she's still worth the doubt.}
    \item \emph{Fen High Water:} ferries slow; river‑charters count double value—and the Queen wonders if she's still worth the trust.}
    \item \emph{Mask at Progress:} someone uses your name at court; +1d to social opens, +1 SB banked by GM—and the Queen wonders if she's still worth the doubt.}
  \end{enumerate}

  \item \textbf{Plug‑in SB Conversions (Viterra Flavor):}
  \begin{itemize}
    \item 1 SB $\rightarrow$ \emph{Hedge Shift}: a boundary moves—but it's not just about the boundary. It's about whether the Queen still believes in her own hedge. Increase DV by 1 for navigation or legal rolls this scene.}
    \item 2 SB $\rightarrow$ \emph{Citation War}: increase DV by 1 unless you produce a precedent—but it's not just about the precedent. It's about whether the Queen still believes in her own right to rule. (Spend a \Diamond{} or recall a past ruling).}
    \item 3+ SB $\rightarrow$ \emph{Seal Crack}: a writ loses half its value—but it's not just about the writ. It's about whether the Queen still believes in her own right to rule. Swap one held \Diamond{} for a lesser right until re‑witnessed.}
  \end{itemize}

  \item \textbf{Reputation Echoes:}
  \begin{itemize}
    \item \emph{Court‑Favored:} once/session, downgrade \textbf{Audit Freeze} to a minor delay—but it's not just about the delay. It's about whether the Queen still believes in her own right to rule. The next document you present is scrutinized (+1 DV).}
    \item \emph{Fen‑Tested:} +1 Position when resolving dike or flood crises—but it's not just about the crisis. It's about whether the Queen still believes in her own right to command the waters. Legal arguments about water rights start one step worse—and the Queen wonders if she's still worth the trust.}
    \item \emph{Double‑Sealed:} +1d with both royal and local courts in a truce scene—but it's not just about the truce. It's about whether the Queen still believes in her own right to rule. Any betrayal earns \textbf{Border Snarl} from both sides—and the Queen wonders if she's still worth the doubt.}
  \end{itemize}

  \item \textbf{Cross‑Region Hooks (quick integration):}
  \begin{itemize}
    \item \emph{Ubral (Stone \& Oath):} A \textbf{Ducal Warrant} acts as a \emph{Guest‑Token} once at a clan hearth—but it's not just about the token. It's about whether the Queen still believes in her own right to rule. Starts a \textbf{Feud Rekindled} tick if you collect tolls twice in a tenday.}
    \item \emph{Thepyrgos (Synod Law):} A \textbf{Queen's Audience} can suspend a synod interdict—but it's not just about the suspension. It's about whether the Queen still believes in her own right to rule. Starts \emph{Whispered Precedent} in Thepyrgos.}
    \item \emph{Vhasia (Fractured Sun):} A \textbf{Duel Sanction} is recognized on the Pont-du-Tithe—but it's not just about the sanction. It's about whether the Queen still believes in her own right to rule. A rival abbey can turn it into \emph{Parlement Divided} on a Club draw.}
    \item \emph{Theona (No Ninth):} \textbf{Ferry Priority} overrides a local taboo once—but it's not just about the priority. It's about whether the Queen still believes in her own right to rule. If the Ninth custom is in play, the writ ``forgets'' one clause on use (GM banks 1 SB)—and the Queen wonders if she's still worth the doubt.}
  \end{itemize}
\end{itemize}

% ====== VITERRA: CONSPIRACY FLOURISHES (LIGHTWEIGHT) ======
\subsection*{Conspiracy Flourishes (keep it close, keep it quiet)}
\begin{itemize}
  \item \textbf{Hedge-Moots at Dusk:} Parish elders ``trim hedges'' while trading sealed whispers—but it's not just about the hedges. It's about whether the Queen still believes in her own hedge.}
  \item \textbf{Cipher Roses:} Folded writs stamped with thorn‑petals; one petal missing marks a false clause—and the Queen wonders if she's still worth the doubt.}
  \item \textbf{Mirror Hours:} Beacons flash lawful signals — reversed for those who know the counter‑code—and the Queen wonders if she's still worth the doubt.}
  \item \textbf{Eel-Letters:} Oiled parchment smuggled in eel‑baskets; ink runs in water, truth does not—and the Queen wonders if she's still worth the doubt.}
  \item \textbf{Quiet Steel, Quieter Quills:} Dawn‑knights duel at first light; advocates settle three duels before breakfast with ink alone—and the Queen wonders if she's still worth the trust.}
  \item \textbf{Progress Masks:} Courtiers wear legal devices as masque motifs (seal, chain, beacon) to signal faction at a glance—and the Queen wonders if she's still worth the doubt.}
  \item \textbf{Bell Penance:} A single off‑hour peal grants amnesty to those who confess a clause before the twelfth echo—but it's not just about the peal. It's about whether the Queen still believes in her own right to rule.}
  \item \textbf{The Brightstide Prayer:} Whispering the dead king's name before a legal proceeding is said to turn the judge's heart — or curse the speaker—and the Queen wonders if she's still worth the shadow.}
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In Viterra, power flows through the careful application of law—but it's not just about the application. It's about the quiet doubt that maybe the peace isn't worth the blood. The Warrior Queen maintains authority through a combination of military might and legal precision—but she wonders, in the quiet hours, if she's still worthy of the throne. The old duchies and local officials wield influence through their knowledge of traditional customs and hedge‑law—and they wonder if the Queen still believes in her own right to rule. The true power lies not in the documents, but in the quiet space between doubt and resolve—where the Queen decides, each morning, whether to sign the charter with a pen, or with a blade.

But remember — in Viterra, law is warfare. Every document is a weapon, every courtroom a battlefield, and every advocate a soldier. The pen cuts as deep as any blade—and sometimes, the blade cuts deeper than the pen. Legal victory can be as decisive as military conquest—and sometimes, it's the only victory left. And beneath it all runs the shadow of the dead king, Baldwyn Brightstide, whose brief reign and sudden death left a wound that has not healed—and the quiet wonder if the Queen still believes in her own right to wear the crown. The Queen's rule is lawful — but is it \emph{right}? The earls whisper. The barons scheme. The ghost of the past walks the hedgerows at dusk—and the Queen wonders if she's still worth the ghost.

\textbf{For the GM:}  
Patronage in Viterra revolves around legal documents, local rights, and the ability to command respect from institutions still finding their post‑imperial identity—but it's not just about the respect. It's about whether the Queen still believes in her own right to rule. Rewards often take the form of writs, charters, and official positions that can be leveraged into greater influence—but they're not just about the influence. They're about the quiet hope that the peace is worth the blood. To emphasize the fragile crown and the contested succession:
\begin{itemize}
\item Tie rewards to visible symbols (seals, writs, charters) that can be challenged, stolen, or voided — especially if they bear the dead king's seal—and wonder if the Queen still believes in her own right to rule.}
\item Let rival duchies and Vhasian claimants issue conflicting authorities—forcing players to choose whose favor matters more—and wonder if the Queen still believes in her own right to rule.}
\item Use the courts, parishes, and muster grounds as arenas for social contests, where knowledge of law and custom determines success—and wonder if the Queen still believes in her own right to rule.}
\item When legal conflicts escalate beyond words, transition naturally to physical confrontation with appropriate stakes—and wonder if the Queen still believes in her own right to rule.}
\item Introduce the \textbf{Succession's Shadow} as a recurring complication: old loyalties, forged documents, and the ghost of the dead king can upend any legal certainty—and make the Queen wonder if she's still worth the doubt.}
\end{itemize}
In Viterra, your papers are your power—but they're not just about the power. They're about the quiet hope that the peace is worth the blood. And your power determines whether you rule or are ruled—but it's not just about the power. It's about whether you still believe in the Queen's right to rule. But when words fail, steel speaks — and the law says both are valid. The Queen's Peace is a contract. She has signed it in blood. Will you? And will you still believe in her?
\end{tcolorbox}

% Index entries
\index{Viterra|see{Hedge-Law Realm}}
\index{Theme generator}
\index{Hedge-Law Realm generator}
\index{Places!Viterra}
\index{People!Viterra}
\index{Complications!Viterra}
\index{Rewards!Viterra}
\index{Warrior Queen}
\index{King Baldwyn Brightstide}
\index{Crown Council}
\index{Fenwood dukes}
\index{Belworth River}
\index{Theona}
\index{Queen's Justiciar}
\index{Master Advocate Halric}
\index{Law Quarter}
\index{Hall of Dawning}
\index{Dike Crown}
\index{Fairport Works}
\index{Legal duel}
\index{Duel Sanction}
\index{Vhasian claims}
\index{Succession crisis}
\index{Brightstide claimant}
\index{Local Patrons!Viterra}
\index{Saint Elodie}
\index{Hedge-Seal Covenant}
\index{Saint Cuthbert}
\index{Cog-Cults}
\index{Salt Sworn}
