\section{The Ways Between --- Obishaal, the Realm of Dreams and Death}
\label{chap:ways-between}

\subsection*{Quick Reference}

\textbf{Tagline:} \textit{Every dream is a door. Every door is a judgment. The dead dream forever; the living borrow their sleep.}

\begin{quote}
  \textbf{The Wayfinder (Elite):}  
  ``I exist only in the moment of choice, the breath between one path and another. Obishaal is not a place—it is a \emph{state}. To guide here is to understand that every dream leads somewhere, but only the virtuous know which somewhere is worth the waking.''
\end{quote}

\begin{quote}
  \textbf{Lost Pilgrim (Commoner):}  
  ``The dead don't leave. They just change addresses. Obishaal is their new neighborhood, and the scales are balanced with virtue. If you're lucky, you'll only visit. If you're not... you'll find yourself climbing stairs that never end, and every landing offers a bargain you can't refuse.''
\end{quote}

\noindent\textbf{Genre / Mood:} Dream‑death metaphysical, virtue‑based ascent, temptation and fall, folk‑surrealism.

\noindent\textbf{Starting Location:} A crossroad shrine of black glass stones, where travelers meet their reflections coming the other way. The shrine is silent except for the soft hum of the stones. A Backwards Pilgrim stands at the crossroads, facing the direction you came from. ``The Thirteenth Stone speaks,'' they say. ``But the path chooses who arrives. What did you bury on the road here? The stones can smell it. And in Obishaal, every buried thing dreams of waking.''

\vspace{2mm}
\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
The Ways Between is a transit way, the eternal crossroads through Obishaal, the element of Dreams and Death, the eighth element after Fire, Life, Air, Fortune, Earth, Fate, Water, Dreams/Death. It is the realm where the settled dead go not to oblivion but to an \emph{ideal afterlife}—a dream that reflects the life they lived. A farmer dreams of endless harvest; a warrior dreams of eternal battle; a merchant dreams of scales that never tip. These dreams are real, and they are paradise or prisons.

The living may enter the Ways through thresholds of sleep, near‑death, or deliberate ritual. Once inside, they find a landscape that shifts with their own psyche—but also with the weight of their choices. \textbf{To ascend} through the layers of Obishaal, a traveler must demonstrate \emph{virtue}: honesty, courage, mercy, self‑sacrifice. \textbf{To fall} is to yield to \emph{vice}: greed, cowardice, cruelty, despair. Each layer is a test, and each test is a dream that tempts you to fail.

At the very bottom, pressed into the silt of the deepest dream, lies \textbf{Ikasha}, She Who Sleeps. Her sisters Inaea and Isoka imprisoned her there eons ago, for Ikasha dreamed of summoning and confronting the Clock Work Monad—a confrontation that, in their judgment, would unmake reality. Now Ikasha dreams fitfully, and her nightmares seep upward, infecting the layers above. Sometimes a traveler who falls far enough can hear her voice, offering escape in exchange for service—a service that would free her.

Between the layers drift the dead who have not yet found their final dream. They are the \emph{Unsettled}, and they will trade memories, secrets, and virtues for a chance to ascend—or to pull you down with them.

\vspace{2mm}
\textbf{What you notice first:} The air tastes of copper and old keys. Your footsteps echo half a beat too late. The waystones are warm—and they pulse like hearts.

\textbf{The one rule that kills newcomers:} Never answer a question asked by a reflection. The answer will bind you to the path you did not choose.
\end{tcolorbox}

\subsection*{Regional Mood: The Dream That Remembers}

Here, the dream is not a thing—it is a witness. It remembers every promise made at a deathbed, every lie told to a dying ear, every virtue you failed to keep. Obishaal does not forgive, but it does not forget either. It offers passage in exchange for truth. The longer you stay, the more the dream learns about you. By the time you reach your destination, you may find that the destination knows you better than you know yourself. Some travelers never leave. They become dream‑echoes, their faces pressed into the glass of a forgotten dream, their names still warm on the lips of the next pilgrim.

\begin{tcolorbox}[colback=black!3,colframe=red!60!black,title={The Ninth Taboo Manifestation}]
On any path in the Ways Between, the ninth step you take in a straight line will land on a threshold you did not see. The wise take eight steps, then pause to look around. In Obishaal, the ninth step is always a dream of falling. Those who fall hear a voice ask, ``What did you owe?'' Answering costs a memory; silence costs a name.
\end{tcolorbox}

% ----------------------------------------------------------------------
% Thin Places — Where the Veil Wears Thin
% ----------------------------------------------------------------------

\subsection*{Thin Places — Where the Veil Wears Thin}

Not every threshold is a door. Some are a \emph{crack}—a place where the world forgot to finish its seam. The Pereshi call them \emph{breath‑holes}, where the mountain's exhalation smells of yesterday. The Tulkani knot a thread and leave it untied, marking where the Hollow's attention wanders. The Aeler do not name them at all. They simply count to eight and leave the ninth breath unspoken.

A \textbf{thin place} is a location where the boundary between the waking world and the Ways Between is unusually permeable. Crossing is not a matter of ritual or intent—it is a matter of \emph{being in the wrong place at the wrong angle}. A wrong step. A miscounted breath. A name spoken at the wrong hour.

\subsubsection*{When Does a Thin Place Open?}
The GM may declare a thin place when:
\begin{itemize}
\item A character rolls a \textbf{critical failure} (all $1$s) on a navigation or survival test in a liminal region (Valewood, Mistlands, high passes, deep mines, ancient ruins).
\item The party camps at a \textbf{crossroads, bridge, or ford} during a new moon, an eclipse, or the ninth hour.
\item A character \textbf{breaks a taboo} (pours the ninth cup, takes the ninth step, speaks the ninth name) in a region where the Hollow is strong.
\item The GM spends \textbf{3+ Story Beats} to introduce a supernatural complication, and the fiction supports a threshold crossing.
\item A player \textbf{willingly seeks} a thin place (rare—most who seek them never return unchanged).
\end{itemize}

\subsubsection*{Signs of a Thin Place (d6)}
Roll or choose when the party enters an area that might be thin:
\begin{enumerate}
\item \textbf{Echoes answer wrong.} Your voice returns half a heartbeat later, or in a different pitch. Sometimes the echo speaks words you did not say.
\item \textbf{Shadows drift.} Your shadow moves when you stand still. Cast by a light source that isn't there.
\item \textbf{The ninth step appears.} A stair, a bridge plank, a stone—where no ninth should exist. Step on it, and the ground shifts.
\item \textbf{Bells ring without clappers.} A distant chime, a temple bell, a levee alarm. The sound comes from a direction you cannot find.
\item \textbf{The air tastes of copper and old keys.} The same taste Saikou Ira describes before every exorcism. The same taste the Aeler name \emph{breath of the deep}.
\item \textbf{A door stands where no door stood.} Grey wood, warm handle, no frame. It does not open for those who knock. It opens for those who do not.
\end{enumerate}

\subsubsection*{Accidental Crossing (Mechanics)}
When a character is in a thin place and the conditions align, the GM may call for a \textbf{Spirit + Resolve test} (DV $4$–$6$, depending on the thinness). On a failure, the character crosses into the Ways Between—not by choice, but by \emph{accident}.

\begin{itemize}
\item \textbf{Partial success ($0 < \text{successes} < \text{DV}$):} The character feels the pull. They may choose to step back (mark $1$ Fatigue) or let themselves cross (gain $1$ Boon, but enter the Ways).
\item \textbf{Miss ($0$ successes):} The character crosses involuntarily. They arrive in a random location in the Ways Between (roll on the table in §\ref{sec:ways-between-thin-destinations}).
\item \textbf{Critical failure (all $1$s):} The character crosses—and something crosses \emph{with} them. A Oath‑Keeper, a Counting Wraith, or worse.
\end{itemize}

Once crossed, the character must navigate the Ways Between using the standard travel procedures (Chapter 46). They cannot simply “turn back”—the threshold that opened may close behind them. Finding the way home requires a \emph{balance}: a story, a memory, or a name offered to the road.

\subsubsection*{Known Thin Places (By Spades Draw)}
The following locations are known thin places, keyed to the \emph{Spades} (place) draws of various regional generators. A GM who draws one of these cards while the party is in that region may declare the area thin, triggering the possibility of an accidental crossing.

\begin{longtable}{p{0.18\textwidth} p{0.32\textwidth} p{0.40\textwidth}}
\caption{Known Thin Places by Region and Spades Rank}\\
\hline
\textbf{Region} & \textbf{Spades Entry (Place)} & \textbf{Thin Place Hazard / Hook} \\
\hline
\endfirsthead
\hline
\textbf{Region} & \textbf{Spades Entry (Place)} & \textbf{Thin Place Hazard / Hook} \\
\hline
\endhead
\hline
\multicolumn{3}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot

Aeler & ♠10 (Reliquary Arcade) & \textit{The Bell‑Crypts.} Bells ring without clappers. A ninth tone pulls listeners into a dream‑echo of the mountain's first collapse. \\
Mistlands & ♠Q (Witchlight Bridge) & \textit{The Ninth Plank.} Step where the plank is missing, and you step into the fog of Obishaal. The Hollow Bride sometimes waits on the far side. \\
Ykrul & ♠A (Sky Steppe) & \textit{The Star‑Road Scar.} Where the grass grows in spirals. Crossing at dusk may lead to the Long Mile of Temptation. \\
Vilikari & ♠A (Foedus Stone) & \textit{The Crack of the Ninth Clause.} Touch the stone's newest crack, and you may hear the Oath‑Keeper's question before you answer. \\
Linn & ♠K (High Jarl's Seat) & \textit{The Beacon That Remembers.} When the beacon burns green, the dead can see you. When it burns blue, they can touch you. \\
Thepyrgos & ♠A (The Unfinished Stair) & \textit{The Ninth Flight.} Climb where no stair should be, and you climb through the fossil stairs of Obishaal. \\
Acasia & ♠A (The Pale Causeway) & \textit{The Causeway of Broken Promises.} Every milestone is a broken oath. Walk it at dusk, and the Walking Scale may find you. \\
Vhasia & ♠K (Sun Palace, Lence) & \textit{The Hall of Shattered Mirrors.} Mirrors show not your face, but a reflection of the path you refused. Step through, and you enter the Crossroads of the Ninth Choice. \\
Viterra & ♠A (Highway Stone) & \textit{The Queen's Milestone.} The stone is cracked. The crack is a door. Those who peer inside see the Bone Corridor of Unfinished Business. \\
Valewood & ♠A (Valeheart Spire) & \textit{The Spire's Shadow.} When the spire phases, its shadow becomes a stair. Climb it, and you enter the Stone Circle of Forgotten Villages. \\
Aelinnel & ♠A (The Green Gate) & \textit{The Gate That Expects Change.} If you cannot pay the toll in truths, the gate may open to the Memory Tunnel of the Dying instead of the forest. \\
Aelaerem & ♠A (Hollow Field) & \textit{The Field of Uncounted Steps.} Step nine times in a straight line, and the Hollow Field becomes the Mist‑Ford of First Doubt. \\
Zakov & ♠A (Serpent's Spine) & \textit{The Reef That Sings.} The harmonic hum is the song of the Deep Drake. Follow it, and you may find the Pressure's lair—or the Thirteenth Stone. \\
Tulkani & ♠A (Ikasha's Dreaming Place) & \textit{The Dreamer's Threshold.} A cave that smells of honey and rust. Entering it is already a crossing. The Wayfinder waits inside. \\
Fhara & ♠A (The First Well) & \textit{The Well of Forgotten Names.} Drink from it, and you may wake in the Waystation of the Virtuous, your name already spent. \\
Sidhi & ♠A (The Sealed Coast) & \textit{The Coast of Unmarked Graves.} The drowned walk here. If you answer a drowned sailor's question, you follow them into the Promise Bridge of the Unsettled. \\
Pereshi & ♠A (The First Keystone) & \textit{The Stone That Asks.} Touch it, and the mountain asks a question. Answer poorly, and you climb the Fossil Stairs of Ascension. \\
Kuvani & ♠K (Khagan's Felt Throne) & \textit{The Throne of Tails.} The saddle has no stirrups. Sit in it, and the Long Mile of Temptation begins beneath you. \\
Ashaan & ♠A (Ikasha's Sleeping Chamber) & \textit{The Sealed Door.} The stone is warm. The lock is a whisper. Those who listen too long cross into the deepest layer—and may never wake. \\
Ngomebe & ♠A (The First City) & \textit{The City That Never Stopped.} A walking city seen only on the horizon. Step onto its shadow, and you step into the Spiral Path of Recursion. \\
Taharka & ♠A (The King's Tomb) & \textit{The Tomb of the Unnamed.} The keystone is a door. The door leads to the Threshold Arch of Virtue. The arch judges. \\
Sekogo & ♠A (The Ukwe's Root‑Chamber) & \textit{The Root That Whispers.} Sit on the throne of roots, and the tree asks one question. If you cannot answer, you fall into the Bone Corridor. \\
Lethai‑ar (Crimson Valley) & ♠A (The First Bridge) & \textit{The Invisible Strand.} Only those who have witnessed nine true oaths can see it. Cross it, and you walk the Promise Bridge of the Unsettled. \\
Theona & ♠A (Lookout Cliff) & \textit{The Cliff of Nine Views.} From here you can see the Green Host riding—and sometimes the Fossil Stairs rising beside them. \\
Oshiira & ♠A (The Ledger Vault) & \textit{The Vault of Unpaid Scales.} The records are warm. Read the ninth page, and the Thirteenth Stone speaks to you. \\
Black River (Thepyrgos–Viterra border) & (Special) & \textit{The Ferryman's Slip.} The river is a threshold. Bryn's boat is not the only one. On moonless nights, the dead row themselves. A coin in the water may buy a crossing—or sell your name. \\
\end{longtable}

\subsubsection*{Denizens of the Ways Between (Expanded)}
The following creatures may appear in the Ways Between, drawn from the bestiaries and expansions. They are not random encounters—they are \emph{judges}. Each has a hunger, a weakness, and a price.

\begin{longtable}{p{0.20\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.30\textwidth}}
\caption{Denizens of the Ways Between}\\
\hline
\textbf{Creature} & \textbf{Source} & \textbf{Hunger} & \textbf{Price to Banish} \\
\hline
\endfirsthead
\hline
\textbf{Creature} & \textbf{Source} & \textbf{Hunger} & \textbf{Price to Banish} \\
\hline
\endhead
\hline
\multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
Counting Wraith & \textit{Witnessed Prey} & Errors (ninth step, ninth breath) & A correct count of eight \\
Oath‑Keeper & \textit{Witnessed Prey} & Unpaid scales & Payment of the original weight \\
Trow & \textit{Witnessed Prey} & Absences, omissions & A memory freely given \\
Neighbor & \textit{Witnessed Prey} & Broken taboos, attention & A never‑told truth (or silence) \\
Poltergeist (Class IV) & \textit{Saikou Ira's Compendium} & Rage, unfinished business & A witnessed apology \\
Deep Drake (Pressure) & \textit{Witnessed Prey / Saikou Ira} & Attention, the ninth breath & A memory of sunlight (or a copper bell) \\
Scale Lich & \textit{Witnessed Prey (appendix)} & Unbalanced judgments & A true name carved in quartz \\
The Hollow Bride & \textit{The Hearth Balance} & Hospitality broken, the ninth cup & An apology witnessed \\
The Walking Scale & \textit{The Salt Balance} & Broken promises & A story never told \\
The Keeper of the Seventh Exhalation & \textit{The Wandering Balance} & The seventh sneeze & A name offered freely \\
\end{longtable}

\subsubsection*{Accidental Crossing Destination (d6)}
\label{sec:ways-between-thin-destinations}
When a character crosses into the Ways Between accidentally, roll to determine where they emerge:
\begin{enumerate}
\item \textbf{The Mist-Ford of First Doubt} (Class III Procession site). Water runs uphill. Footsteps echo backwards. The dead march here at midnight.
\item \textbf{The Bone Corridor of Unfinished Business} (Anchor graveyard). Shadows walk the walls but never touch the floor. A Poltergeist may be forming.
\item \textbf{The Crossroads of the Ninth Choice} (Threshold nexus). All four paths lead to the same destination—but different \emph{yesterdays}. A Trow sits at the center, smoking.
\item \textbf{The Promise Bridge of the Unsettled} (Oath‑Keeper territory). Cross with a vow, and the bridge remembers. The toll is a story you have never told.
\item \textbf{The Long Mile of Temptation} (Vice's testing ground). The road stretches differently for each traveler. A Walking Scale appears at the halfway mark.
\item \textbf{The Thirteenth Stone (Ikasha's Seal)}. The path ends at a warm stone. The stone whispers. Beneath it, She Who Sleeps dreams of freedom. A Deep Drake may be coiled nearby.
\end{enumerate}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={GM Guidance: Thin Places as Pressure, Not Punishment}]
Thin places are not traps. They are \emph{complications}. A party that accidentally crosses into the Ways Between should not feel that they have “failed.” They have simply entered a different kind of scene—one where negotiation matters more than steel.

\begin{itemize}
\item \textbf{Offer a way out.} The Ways Between always offer a path home. The path may demand a toll (a memory, a story, a name), but it exists.
\item \textbf{Use the travel rules.} Treat the Ways Between as a leg of a journey. Draw cards. Set timers. The destination is “home” or “the next threshold.”
\item \textbf{Let players use their bonds.} A character who calls on a bond while lost in the Ways may hear a familiar voice—or see a familiar face—guiding them. The road remembers who weighs what.
\item \textbf{The ninth is always watching.} In any scene set in the Ways Between, the ninth card drawn (or the ninth roll made) triggers a manifestation of the Hollow. Not a fight—a \emph{question}. The players must answer truthfully, or the road takes a memory.
\end{itemize}
\end{tcolorbox}

\vspace{2mm}
\noindent\textit{Set down by the Grey Wanderer, after a night spent in a thin place near the Black River. The fire went out three times. The shadows did not. The ninth bell rang once—and then was silent.}

\noindent\textit{The Ways Between are not a dungeon. They are a conversation. Speak carefully. The road listens.}

\section*{Spades --- Places (Dream Layers)}
\label{sec:ways-between-places}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Mist-Ford of First Doubt} --- Water runs uphill; footsteps echo backwards through time. `[THRESHOLD]`
  \hfill \textit{The ford shows you wading through your own past. Do not look down at the water.}
\item \textbf{Bone Corridor of Unfinished Business} --- Shadows walk the walls but never touch the floor; light as memory. `[DEATH]`
  \hfill \textit{The bones are from those who chose vice and fell. They shift when you blink.}
\item \textbf{Threshold Arch of Virtue} --- Shows a reflection as someone you might have been; the past as a mirror. `[TRUTH]`
  \hfill \textit{The reflection has a scar you never got. It asks how you avoided it.}
\item \textbf{Spiral Path of Recursion} --- Ascends but returns to the same stone marker; the journey as meditation. `[MAGIC]`
  \hfill \textit{Each loop, the stone moves closer to the center. Or you do.}
\item \textbf{Promise Bridge of the Unsettled} --- Cross with a vow and it remembers your weight; an oath as a toll. `[OATH]`
  \hfill \textit{The bridge's planks are old letters. Read one, and you may owe the writer.}
\item \textbf{Crossroads of the Ninth Choice} --- All four paths lead to the same destination, but different tomorrows. `[THRESHOLD]`
  \hfill \textit{The center stone is warm. A coin placed there will vanish—or be accepted.}
\item \textbf{Stone Circle of Forgotten Villages} --- Marks where a village used to be; the houses are now constellations. `[DEATH]`
  \hfill \textit{The stars are too close. You can touch them. They crumble like ash.}
\item \textbf{Memory Tunnel of the Dying} --- Hear the thoughts of those who passed here; the past as an echo. `[MEMORY]`
  \hfill \textit{The thoughts are in a language you almost understand. One word is your name.}
\item \textbf{Fossil Stairs of Ascension} --- Each step is an eye that watches your ascent; the climb as scrutiny. `[FEAR]`
  \hfill \textit{The eyes belong to those who fell here. They blink in sequence.}
\item[J] \textbf{Waystation of the Virtuous} --- The keeper trades in unfinished conversations; words as currency. `[SOCIAL]`
  \hfill \textit{The keeper has no face. It wears the faces of those who paused too long.}
\item[Q] \textbf{Junction of Vice and Virtue} --- Paths physically collide; you must choose which reality to follow. `[THRESHOLD]`
  \hfill \textit{The collision sounds like breaking glass. The glass reforms into a door.}
\item[K] \textbf{Long Mile of Temptation} --- The road stretches differently for each traveler; distance as perception. `[MAGIC]`
  \hfill \textit{You are the only one walking. The others are already at the end.}
\item[A] \textbf{The Thirteenth Stone --- Ikasha's Seal} --- Where the path reveals its true destination; truth as revelation. `[TRUTH]`
  \hfill \textit{The stone whispers. The whisper is a promise you made to yourself and broke. Beneath it, the stone is warm—and dreaming.}
\end{enumerate}

\paragraph*{(Wayfarer/Spirit/Dream-walker)} Lost pilgrim who thinks they're going home but forgot where; toll‑taker accepting payment in forgotten memories.

\section*{Hearts --- Travelers \& Guides}
\label{sec:ways-between-people}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Lost Pilgrim} --- Thinks they're going home but forgot where; destination as a memory. `[TRAVEL]`
  \hfill \textit{Their pack is full of stones. Each stone is a forgotten promise.}
\item \textbf{Toll-Taker (Virtue Assessor)} --- Accepts payment in virtues you didn't know you had. `[SOCIAL]`
  \hfill \textit{Their scales weigh truth and echo. A balanced scale means you pass—for now.}
\item \textbf{Wayward Guide} --- Directions technically correct, morally questionable. `[TRAVEL]`
  \hfill \textit{They point with a broken compass. The needle spins, but the path is true.}
\item \textbf{Dream-Merchant} --- Carries a sack of nightmares and one perfect dream for sale. `[MAGIC]`
  \hfill \textit{The perfect dream is their own. They are selling it because they cannot wake up.}
\item \textbf{Child-Ghost of Unfinished Life} --- Knows shortcuts but charges in riddles; wisdom as a puzzle. `[DEATH]`
  \hfill \textit{The riddles have no answer—only a choice. Choose wrong, and you become a child.}
\item \textbf{Wounded Walker (Vice-Bound)} --- Bleeding metaphor; their pain grows thorns along the path. `[FEAR]`
  \hfill \textit{The thorns have faces. Each face is someone they failed to save.}
\item \textbf{Map-Merchant of Impossible Dreams} --- Selling maps to places that don't exist yet; the future as a commodity. `[TRUTH]`
  \hfill \textit{The maps change as you watch. The destination is always one page ahead.}
\item \textbf{Guide-Dog (Ancestral Spirit)} --- Made of shadow and starlight; follows those who walk with purpose. `[SPIRIT]`
  \hfill \textit{It never barks. It whines when you are about to make a mistake.}
\item \textbf{Backwards Pilgrim (Dream-Echo)} --- Moving through time; their steps erase what just happened. `[MEMORY]`
  \hfill \textit{They are walking toward their birth. Every step, they forget something.}
\item[J] \textbf{Soul-Ferryman} --- Carries the luggage of the living; a burden as a service. `[DEATH]`
  \hfill \textit{The luggage is full of things you never said. The ferryman knows them all.}
\item[Q] \textbf{Road's Child (Dream-Kin)} --- Born where paths crossed, raised by waymarks; navigation as instinct. `[TRAVEL]`
  \hfill \textit{They have no navel. They were never born—they simply began.}
\item[K] \textbf{Dead-Road Keeper} --- Ensures proper passage for those who should not walk. `[DEATH]`
  \hfill \textit{Their lantern burns a cold blue. The flame does not flicker.}
\item[A] \textbf{The Wayfinder (Dream Avatar)} --- Exists only in the moment a choice is made; a decision as an entity. `[LEGEND]`
  \hfill \textit{They are every guide you ever needed. They are also every road you refused.}
\end{enumerate}

\paragraph*{(Veil‑Thin/Dream‑Bleed/Wayward)} Path loops back to show your funeral preparations; reality thins---you see the dreams of sleeping travelers; Ikasha's nightmare bleeds into the upper layers.

\section*{Clubs --- Complications/Threats}
\label{sec:ways-between-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Funeral Loop} --- The path shows your funeral preparations; death as a destination. `[DEATH]` `[FEAR]`
  \hfill \textit{The mourners are people you have not met yet. They seem to know you well.}
\item \textbf{Reality Thins (Ikasha's Dream Bleed)} --- You see the dreams of sleeping travelers; the boundary as a veil. `[MAGIC]`
  \hfill \textit{A dreamer mutters your name. They are dreaming of a choice you will make tomorrow.}
\item \textbf{Waymark Lies (Vice's Test)} --- Points the wrong way; trust leads to places that never were. `[TRUTH]`
  \hfill \textit{The waymark has a crack. Through the crack, you see a path you should have taken.}
\item \textbf{Dream-Bleed} --- Waking memories become someone else's nightmares. `[MEMORY]` `[FEAR]`
  \hfill \textit{You remember a childhood fear that was not yours. It still terrifies you.}
\item \textbf{Toll Demanded (Virtue Payment)} --- Currency you did not know you carried; payment as a revelation. `[SOCIAL]` `[WEIGHT]`
  \hfill \textit{The toll is a secret you told a dead person. They are now standing behind you.}
\item \textbf{Path Split (Virtue vs. Vice)} --- Each version of you remembers differently; identity as a choice. `[MAGIC]`
  \hfill \textit{One you is smiling. One you is weeping. One you is already walking away.}
\item \textbf{Gravity Shifts (Guilt's Weight)} --- Emotional weight grows physically heavy; a burden as physics. `[WEATHER]`
  \hfill \textit{Your pack feels heavier. Inside is a regret you forgot you carried.}
\item \textbf{Time-Sickness (Dream Chronology)} --- You arrive before you left; paradox shadows follow. `[MAGIC]`
  \hfill \textit{Your past self waves at you. Do not wave back.}
\item \textbf{Truth Payment (Ikasha's Price)} --- The road remembers your lies and demands truth as payment. `[TRUTH]` `[WEIGHT]`
  \hfill \textit{The lie is carved on a stone. The stone is warm. It knows your temperature.}
\item[J] \textbf{Crossroads Judgment (Virtue Audit)} --- The path you chose judges the ones you did not; decision as karma. `[TRUTH]` `[SOCIAL]`
  \hfill \textit{A ghost of the road not taken asks why. You have one sentence to answer.}
\item[Q] \textbf{Memory-Thief (Unsettled Dead)} --- A rest stop keeper pays in counterfeit recollections. `[MEMORY]` `[SOCIAL]`
  \hfill \textit{The memory is almost right. The color of the sky is wrong.}
\item[K] \textbf{Forbidden Path (Ikasha's Descent)} --- Opens only for those already lost; loss as a key. `[DEATH]`
  \hfill \textit{The path smells of old tears. You taste salt on your lips.}
\item[A] \textbf{Convergence (Ikasha's Waking)} --- All travelers arrive together, none recall how; arrival as a mystery. `[MAGIC]` `[LEGEND]`
  \hfill \textit{You recognize everyone. You have never seen any of them before. The air tastes of honey and rust. Ikasha is dreaming of you.}
\end{enumerate}

\paragraph*{(True Name/Safe Passage/Waywisdom)} Waymark that always points toward your next important choice; a token of passage that the road recognizes as belonging.

\section*{Diamonds --- Rewards/Leverage}
\label{sec:ways-between-rewards}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Choice Compass (Virtue's Needle)} --- Always points toward your next important choice. `[TRAVEL]`
  \hfill \textit{The needle is a splinter of the Thirteenth Stone. It hums when you are near a decision.}
\item \textbf{Passage Token (Dream-Marked)} --- The road recognizes you as one who belongs. `[SOCIAL]`
  \hfill \textit{The token is a braid of your own hair. You do not remember cutting it.}
\item \textbf{Dream-Catcher} --- Filters nightmares from your rest; sleep as a sanctuary. `[MAGIC]`
  \hfill \textit{The catcher is woven from old promises. It smells of lavender and regret.}
\item \textbf{Truth-Compass (Ikasha's Fragment)} --- Points to what you most need to know, not what you want. `[TRUTH]`
  \hfill \textit{The compass has no needle. The housing spins until you speak a truth.}
\item \textbf{Memory-Anchor} --- Keeps you from losing yourself on deeper paths. `[MEMORY]`
  \hfill \textit{The anchor is a river stone. Your name is carved on it in a language you have never seen.}
\item \textbf{Safe Haven (Virtue's Rest)} --- A guaranteed rest stop without complication. `[TRAVEL]`
  \hfill \textit{The haven is a dry arch. The arch is a doorframe with no door.}
\item \textbf{Guide-Light (Virtue's Beacon)} --- Burns the color of your truest intention; purpose as a beacon. `[SOCIAL]`
  \hfill \textit{The light is blue. Blue means you are honest. Red means you are not.}
\item \textbf{Path-Cutter (Dream Shears)} --- Cuts distance through metaphor, not space. `[MAGIC]`
  \hfill \textit{The cutter is a pair of shears. Cut a memory, and the road between here and there vanishes.}
\item \textbf{Scale-Clearing (Virtue's Mercy)} --- The road forgives one weight you thought you carried. `[WEIGHT]`
  \hfill \textit{The weight was owed to a dead person. They are no longer dead on this path.}
\item[J] \textbf{Crossroads Boon} --- Choose among three paths, each exactly where you need to be. `[LEGEND]`
  \hfill \textit{The paths are labeled: Regret, Forgive, Forget. Choose carefully.}
\item[Q] \textbf{Way-Wisdom (Dream-Logic)} --- Understand the language of paths and signs for one journey. `[TRAVEL]`
  \hfill \textit{The waystones whisper. They are complaining about the last traveler. It was you.}
\item[K] \textbf{Grace Passage (Virtue's Shield)} --- Walk safely through any dangerous crossing for one night. `[TRAVEL]`
  \hfill \textit{The grace is a bell. Ring it, and the road will pretend not to see you.}
\item[A] \textbf{Road's Name (Ikasha's True Name)} --- Call the path by its true name and command its nature. `[LEGEND]`
  \hfill \textit{The name is a single syllable. Speaking it will change the road—and you.}
\end{enumerate}

\section*{Quick use notes}
\label{sec:ways-between-quick-use}
\begin{itemize}
\item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Timer ($2$–$5 \rightarrow 4$, $6$–$10 \rightarrow 6$, J/Q/K $\rightarrow 8$, A $\rightarrow 10$).
\item Diamonds are codified outcomes (true names/safe passages/waywisdom) that change position rather than call for a roll.
\item If any A appears, echo \textbf{dream‑road} motifs—reflections that lie, paths that judge, destinations that choose the traveler. Also, the presence of an Ace hints that Ikasha's dream is particularly close.
\end{itemize}

\subsection*{Additional Features}
\label{sec:ways-between-features}
\begin{itemize}
\item \textbf{Dream-Logic Navigation:} Any Ace introduces a metaphysical requirement (a sacrifice, a taboo, an exchange) that must be satisfied to pass. Name it at the table; the road will enforce it.
\item \textbf{Reflection Points:} Face cards (J, Q, K) show alternate selves or unlived choices. Treat them as temporary NPCs or scene tags that can help, hinder, or tempt.
\item \textbf{Memory Currency:} Diamonds may be traded for knowledge or safe passage as if they were memories. Describe the memory paid; the path (or its keeper) pays in kind.
\item \textbf{Virtue and Vice Timers:} Each character in Obishaal tracks a small personal \textbf{Virtue} [3] and \textbf{Vice} [3] timer. Virtue advances when they act with honesty, courage, mercy, or self‑sacrifice. Vice advances when they act with greed, cowardice, cruelty, or despair. When one timer fills, the character \textbf{ascends} (Virtue) to a higher layer of Obishaal or \textbf{falls} (Vice) to a lower one. Ascension offers a boon; falling offers a temptation (a cursed gift from the layer's vice‑spirit). The timers reset after each layer transition. Reaching the bottom (Vice filled in the deepest layer) allows Ikasha to speak directly to the character.
\item \textbf{Ikasha's Whisper (Patron Intervention):} When a character falls to the lowest layer (or deliberately seeks her out), they may hear Ikasha's voice. She offers a \emph{Fragment of Freedom}: $+2$ dice to any roll involving escape, secrecy, or breaking bonds, but each use marks $1$ \textbf{Obligation to Ikasha} and advances a \textbf{Waking Dream} [6] timer. When that timer fills, Ikasha uses the character as a vessel to attempt an escape—a campaign‑level event.
\item \textbf{Local Patrons (Syncretic Names):}
  \begin{itemize}
    \item \textbf{The Unnamed Guide} — The Traveler. \textit{The road that appears only when you are lost.}
    \item \textbf{The Mirror Listener} — The Witness. \textit{The reflection that tells you what you need to hear, never what you want.}
    \item \textbf{The Threshold Child} — The Sealed Gate. \textit{The one who was never born, who guards the places where paths end.}
    \item \textbf{The Echo King} — The Ninth. \textit{The voice that answers when you speak to the empty road.}
    \item \textbf{The Dreamer Below} — Ikasha (imprisoned). \textit{She who sleeps and dreams of waking. Her whispers are bargains; her freedom is a promise.}
  \end{itemize}
\end{itemize}

\subsection*{Lore Echoes (Cross‑Expansion Hooks)}
\begin{itemize}
  \item \textbf{The Bitter Root:} Daughters of the Cord know the Ways Between as the Cord's shadow. A lost Hollowed wanders here, its thread trailing into the mist. Follow it, and you may find a Sister who cut herself free—and who now tends Ikasha's dream.
  \item \textbf{The Threshold Compendium:} Saikou Ira's silver mirror shows the Ways Between as a tangled knot of unfulfilled balances. He has walked three paths here. He will not speak of the fourth, where he glimpsed Ikasha's face.
  \item \textbf{Malachai, the Chained Angel:} A silver coin that never spends lies at every crossroads. Pick it up, and the road will offer you a shortcut—but the coin will reappear in your pocket at the next junction, heavier. In Obishaal, the coin is warm and pulses like a heartbeat.
  \item \textbf{The Grey Wanderer's Grimoire:} Psions call the Ways Between ``the Stair.'' They climb it in their dreams, trading memories for glimpses of paths not yet walked. Some never wake. The Grey Wanderer claims that the Stair is Ikasha's spine.
\end{itemize}

\subsection*{Ways Between Superstitions (burn for $+1$ die once)}
\begin{itemize}
  \item \textbf{Bread to the Stones:} Crumble a crust at the base of a waystone. Ignore the first \emph{Waymark Lies} penalty. `[SUPERSTITION]`
  \item \textbf{Knot the Path:} Tie a single knot in a piece of string and hang it from a milestone. Cancel \emph{Path Split} or take $--1$ die on your next navigation roll (your choice). `[SUPERSTITION]`
  \item \textbf{Name to the Crossroads:} Whisper a dead traveler's name into the center stone. Gain $+1$ Effect when bargaining with a Toll‑Taker this scene, but mark \emph{Obligation} to the Wayfinder. `[SUPERSTITION]`
\end{itemize}

\subsection*{Thematic SB Spend Table}
\label{sec:ways-between-sb}

\paragraph{Minor Complications (1 SB)}
\begin{itemize}
\item \textbf{Exposure:} Your actions draw unwanted attention from \textbf{path keepers or wayward spirits}.
\item \textbf{Noise:} Sounds of your actions alert nearby \textbf{travelers or memory echoes}.
\item \textbf{Trace:} Evidence of your passage marks your route for \textbf{trackers or reflection points}.
\item \textbf{Delay:} A brief but meaningful setback costs you \textbf{time or favorable path alignment}.
\item \textbf{Supply Strain:} Mark $+1$ segment on a relevant \textbf{resource timer}.
\end{itemize}

\paragraph{Moderate Setbacks (2 SB)}
\begin{itemize}
\item \textbf{Alarm Raised:} \textbf{Toll‑taker or wayward guide} becomes aware and begins responding.
\item \textbf{Position Lost:} You lose advantageous ground/cover/stealth due to \textbf{path shift or reality thinning}.
\item \textbf{Foe Appears:} A \textbf{memory‑thief or forbidden path entity} arrives on scene.
\item \textbf{Gear Trouble:} A piece of equipment becomes \textbf{Compromised/Neglected}.
\item \textbf{Lock/Barrier:} A simple obstacle now requires a test to overcome.
\end{itemize}

\paragraph{Serious Trouble (3 SB)}
\begin{itemize}
\item \textbf{Reinforcements:} Additional \textbf{wayward spirits, path entities, or alternate selves} arrive.
\item \textbf{Key Gear Breaks:} A crucial tool/weapon becomes temporarily unusable.
\item \textbf{Major Twist:} The situation fundamentally changes - \textbf{path loops/time shifts/reality fractures}.
\item \textbf{Rail Tick:} Advance a relevant campaign/front timer by $1$ segment.
\item \textbf{Condition Applied:} Mark \textbf{Fatigue 1/Harm 1/Condition} appropriate to fiction.
\end{itemize}

\paragraph{Major Turns (4+ SB)}
\begin{itemize}
\item \textbf{Trap Springs:} A prepared danger activates with full effect.
\item \textbf{Authority Arrival:} \textbf{The Wayfinder, Dead‑Road Keeper, or Road's Child} intervenes.
\item \textbf{Scene Shift:} The environment changes dramatically - \textbf{reality thins/paths converge/time loops}.
\item \textbf{Patron Omen:} Divine/arcane forces take notice - \textbf{omen appears/blessing lost/curse manifests}.
\item \textbf{Narrative Pivot:} The story takes an unexpected turn that reframes objectives.
\end{itemize}

\paragraph{Region-Specific SB Options}
\begin{itemize}
\item \textbf{Obishaal (Dream‑Logic):} Paths rearrange without warning, reflections speak unbidden, destinations shift mid‑journey. A character may hear Ikasha's whisper as a free omen (the GM may spend $1$ SB to make it a false omen).
\item \textbf{Obishaal (Memory Currency):} Memories become tangible, forgotten knowledge resurfaces, past and present blur. Trading a memory here may also tick the \emph{Virtue} or \emph{Vice} timer depending on the memory's nature.
\item \textbf{Obishaal (Virtue and Vice):} Decisions echo forward, alternate selves appear, paths judge moral weight. The GM may spend SB to force a character to test a virtue (e.g., ``Resist the temptation to lie'') or to offer a vice that ticks their Vice timer.
\end{itemize}

\subsection*{Path‑Moves \& Procedures (Optional Depth)}
\label{sec:ways-between-procedures}

\paragraph{Weighing of Steps (Travel Move)}
When the company enters a liminal path (Mist‑Ford, Bone Corridor, etc.), choose one:
\begin{itemize}
  \item \textbf{Pay the Toll:} Each PC names a small truth or a fond memory (GM: $1$ SB banked, $+1$ Position on navigation).
  \item \textbf{Refuse the Toll:} $+1$d on Speed, but start a \textbf{Way‑Scale} [4] timer.
  \item \textbf{Bargain the Toll:} Trade a \emph{Diamond} reward as if it were a memory (see Memory Exchange below); cancel one imminent \emph{Club} card in this scene.
\end{itemize}
When \textbf{Way‑Scale} fills, trigger \emph{Crossroads Judgment} at the next junction.

\paragraph{Reflection Duel (Social/Moral Test)}
When confronted by an alternate self or \emph{Reflection Point}, state what that self wants from you and roll your best social action.
\begin{itemize}
  \item \textbf{Success:} Gain a \emph{Guide‑Light} tag for the scene and clear $1$ tick from \textbf{Paths Decide}.
  \item \textbf{Mixed:} Trade places for one exchange or take a \textbf{Condition: Disoriented}.
  \item \textbf{Miss:} The reflection walks away with one of your \emph{Diamonds} (GM picks the most narratively apt).
\end{itemize}

\paragraph{Memory Exchange (Rates \& Risks)}
\begin{tabular}{p{0.36\textwidth} p{0.28\textwidth} p{0.30\textwidth}}
\textbf{Memory Offered} & \textbf{Buys You} & \textbf{Side Effect} \\
\hline
A childhood smell, a forgotten nickname & $+1$ Position on one crossing & A familiar place feels strange for a scene \\
A first failure or small betrayal & Cancel one minor Complication & You gain a \emph{tell} (the Ways recognize you) \\
A true fear you have never spoken & \emph{Guide‑Light} or \emph{Way‑Cord}‑like effect & A dream pursues you for $1$–$3$ nights \\
The details of a kept promise & \emph{Safe Passage} for the party & An NPC tied to that promise misremembers you \\
The meaning behind a scar & Convert one \emph{Club} into \emph{Trace} only & The scar “moves” (cosmetic) until the next dawn \\
A perfect day (all details) & Treat the next \emph{Heart} face as an ally & Start \textbf{Dream‑Bleed} [3] on the map \\
\end{tabular}

\medskip
\textbf{Exchange Rule:} If you pay the same \emph{kind} of memory twice in a journey, start \textbf{Hollowing} [4]. When it fills, you gain the \emph{Backwards Pilgrim} tag until you reclaim what was lost.

\subsection*{Way‑Keys \& Taboos (d6)}
\begin{enumerate}
\item \textbf{Speak Last}: Only the final word said at a gate is true; others are “practice.” First speakers take $--1$d on their next test.
\item \textbf{Step Odd}: Crossings require an odd number of steps; even counts loop you back one scene.
\item \textbf{Name Nothing}: True names wake \emph{Name‑Theft}; use titles or lose a \emph{Diamond}.
\item \textbf{Wet Iron}: A drop of water on iron cancels one glamour once per scene.
\item \textbf{Back Gift}: Leave something behind to go forward; refuse and tick \textbf{Way‑Scale}.
\item \textbf{Answer Bent}: Answer a waystone's question with a metaphor, not a fact, or trigger \emph{Truth Payment}.
\end{enumerate}

\subsection*{Boss Seeds (three‑phase foes)}
\begin{itemize}
\item \textbf{The Archivist of Unmade Roads (cloak stitched from surveyor's cords)} --- \emph{Phases}: Receipts of Refusal $\rightarrow$ Stay of Journey $\rightarrow$ Unfiled Options. \emph{Cracks}: a \textbf{Choice Compass} reveals the true path; a \textbf{Scale‑Clearing} voids his claim.
\item \textbf{The Toll‑Matriarch Beneath the Bridge (many hands, one voice like wet stone)} --- \emph{Phases}: Truth Payment $\rightarrow$ Scale‑Clearing $\rightarrow$ Forbidden Path. \emph{Cracks}: a \textbf{Passage Token} marks you as belonging; a \textbf{Road's Name} commands her to stand aside.
\item \textbf{The Wayfinder's Shadow (your silhouette, two steps ahead)} --- \emph{Phases}: Crossroads Judgment $\rightarrow$ Guide‑Light $\rightarrow$ Haste Mark. \emph{Cracks}: a \textbf{Memory‑Anchor} keeps you grounded; a \textbf{Grace Passage} slips past its gaze.
\item \textbf{Ikasha's Nightmare (Dream-Eater)} --- Appears only in the deepest layer. \emph{Phases}: Whispers $\rightarrow$ Temptation $\rightarrow$ Consumption. \emph{Cracks}: a \textbf{Virtue} filled timer can be spent to reject the nightmare; a memory of true self‑sacrifice (cost $1$ permanent Bond) banishes it entirely.
\end{itemize}

\subsection*{Encounter Seeds (quick hooks)}
\begin{itemize}
\item \textbf{The Promise Bridge's Receipt:} An old vow you never spoke was logged here. To cross, define when you would have sworn it---and to whom.
\item \textbf{Thirteenth Stone Auction:} The stone sells three possible endings. Each bidder is you from a different failure.
\item \textbf{Bone Corridor Choir:} Shadows sing a name wrong by one letter. Correct it and gain \emph{Truth‑Compass}; ignore it and start \textbf{Name‑Theft}.
\item \textbf{Ikasha's Finger Bone:} A character finds a bone that pulses with dream‑light. If they hold it while sleeping, they can speak to Ikasha directly—but each time, the \textbf{Waking Dream} timer ticks.
\end{itemize}

% Extensions (Plug-in)
\subsection*{Extensions (Plug-in)}
\label{sec:ways-between-extensions}

\paragraph{Choice Dial (Resolve $\leftrightarrow$ Drift)}
Track the party's footing between firm intent and wandering.
\begin{itemize}
  \item \emph{Resolve High:} $+1$ Position on \emph{Path Split} tests; the first \emph{Dream‑Bleed} each session hits harder (start at $2$).
  \item \emph{Drift High:} $+1$ Effect when bargaining with way‑entities; the first \emph{Truth Payment} each session costs a cherished \emph{detail}.
  \item Center the Dial by naming aloud the next choice \emph{and} why you might refuse it; clear $1$ tick from \textbf{Way‑Scale}.
\end{itemize}

\paragraph{Counter‑Seal (once/scene, if justified)}
Break a single metaphysical instruction (taboo, key, or omen) by presenting a memory in symbol form (a token, a letter, a sketch). Gain $+1$d on that crossing; the GM banks $1$ SB as a \emph{Marked Consequence} to spend when reflections return.

\paragraph{Start‑of‑Scene Murmurs (1d6)}
\begin{enumerate}
  \item \emph{Second Sunrise:} light arrives twice; your first missed navigation roll becomes a mixed success.
  \item \emph{Name Draft:} the air steals a syllable; your next lie sounds true once.
  \item \emph{Echo Scale:} your footprints owe a step—skip a square or tick \textbf{Way‑Scale} $+1$.
  \item \emph{Pale Bell:} time hiccups; the first \emph{Club} can be delayed until the end of the scene.
  \item \emph{Traveler's Chorus:} distant voices agree with you; $+1$d on appeals to path entities.
  \item \emph{Hollow Weather:} rain that does not wet; conditions look dire but impose no penalties this scene.
\end{enumerate}

\paragraph{Plug‑in SB Conversions (Obishaal Flavor)}
\begin{itemize}
  \item $1$ SB $\rightarrow$ \emph{Footprint Vote}: your tracks choose a different direction; increase DV by $1$ on any roll that relies on following a trail.
  \item $2$ SB $\rightarrow$ \emph{Echo of Refusal}: a choice you avoided manifests as a minor obstacle (a locked gate, a forgotten toll). If the choice involved a vice, also tick the Vice timer of the character who refused.
  \item $3+$ SB $\rightarrow$ \emph{Threshold Cracks}: the path fractures; gain a temporary \textbf{Path‑Cutter} effect but start \textbf{Convergence} [4] (all travelers you have met will arrive together at the next junction). Additionally, Ikasha's dream seeps through the cracks; each character hears a single word of her voice.
\end{itemize}

\paragraph{Cross‑Regional Conversions (Using Diamonds Across the Map)}
\begin{itemize}
  \item \emph{Valewood:} \textbf{Road's Name} may count as a \emph{Valeheart Clause} once if spoken under leaves; doing so creates \textbf{Court Tithe} [2].
  \item \emph{Mistlands:} \textbf{Grace Passage} functions as \emph{Pall Indulgence} over one levee crown at night; ring a bell afterward or start \textbf{Witchlight Count}.
  \item \emph{Vhasia:} \textbf{Way‑Wisdom} can be presented as “pilgrim right” to bypass one \emph{Interdict} barrier; begin a \textbf{Parlement Divided} tick.
  \item \emph{Viterra:} \textbf{Passage Token} can be notarized into a temporary \emph{Ferry Priority}; on use, advance \textbf{Border‑Lace}.
  \item \emph{Ubral:} \textbf{Memory‑Anchor} counts as a \emph{Guest‑Token} at one hearth if you recount a family road‑tale.
  \item \emph{Silkstrand:} \textbf{Path‑Cutter} doubles as a night‑boat route once; spend $1$ favor with a smuggler or trigger \emph{Seal Counterfeit}.
  \item \emph{Linn:} \textbf{Choice Compass} acts as a \emph{Pilot's Token} in a reed‑maze if you pay a name to the water.
  \item \emph{Theona:} \textbf{Crossroads Boon} allows counting “eight‑and‑one” to cross an \emph{Uncounted Bridge} safely once.
\end{itemize}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Patronage \& Power}]
In Obishaal, power flows through understanding the metaphysical nature of dreams, the ability to navigate virtue and vice, and the careful management of one's weight before the scales. The true authorities are those who know the True Names of dream‑layers, can mediate between the living and the Unsettled, and understand the currency of virtue.

\textbf{For the GM:}  
Patronage in Obishaal revolves around guidance, safe passage, and the ability to navigate moral tests. Rewards often take the form of dream‑tokens, fragments of Ikasha's memory, or knowledge of a layer's True Name. To emphasize this:
\begin{itemize}
\item Tie rewards to visible symbols (compasses, anchors, bells) that can be challenged, stolen, or voided by the dream itself.
\item Let rival guides (Virtue vs. Vice) issue conflicting directions, forcing players to choose whose favor matters more.
\item Use the crossroads, waystations, and dream‑layers as arenas for moral contests, where a single act of mercy can ascend you—and a single act of cruelty can damn you.
\end{itemize}
In Obishaal, your virtue is your power, and your power determines whether you wake or sleep forever.
\end{tcolorbox}

% Index entries
\index{Ways Between|see{Obishaal}}
\index{Obishaal}
\index{Ikasha}
\index{Dreams and Death}
\index{Virtue and Vice}
\index{Theme generator}
\index{Spiritways and Veilways generator}
\index{Places!Obishaal}
\index{People!Obishaal}
\index{Complications!Obishaal}
\index{Rewards!Obishaal}
\index{The Wayfinder}
\index{Dead-Road Keeper}
\index{Road's Child}
\index{Soul-Ferryman}
\index{Toll-Taker}
\index{Thirteenth Stone}
\index{Long Mile}
\index{Promise Bridge}
\index{Crossroads}
\index{Memory Tunnel}
\index{Fossil Stairs}
\index{Choice Compass}
\index{Passage Token}
\index{Dream-Catcher}
\index{Truth-Compass}
\index{Memory-Anchor}
\index{Road's Name}
\index{Grace Passage}
\index{Local Patrons!Obishaal}