% !TEX root = ../fates_edge_players_guide.tex

\chapter{Quick Reference Sheets}

These condensed sheets summarize the core procedures, costs, and tables for use at the table. They are designed for GMs and players to keep the flow of play without flipping through sections.

\section{Core Mechanic: The Art of Consequence}

\subsection*{1. Approach}
Player states intent and method (Attribute + Skill).

\subsection*{2. Execution}
Roll dice pool of d10s:
\begin{itemize}
  \item Each \textbf{6+} = 1 Success
  \item Each \textbf{1} = 1 Complication Point (CP)
\end{itemize}

\subsection*{3. Outcome}
Compare Successes to Difficulty Value (DV):

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Outcome} & \textbf{Effect} \\
\midrule
Clean Success & Intent achieved crisply (DV met, no CP) \\
Success \& Cost & Intent achieved; GM spends CP for complications \\
Partial & Progress with fork (accept cost OR concede ground) \\
Miss & No progress; GM spends CP OR offers Devil's Bargain \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Description Ladder}
\begin{description}
  \item[Basic] Roll as-is; tally Complications.
  \item[Detailed] Re-roll one 1.
  \item[Intricate] Re-roll all 1s; add one positive flourish if successful.
\end{description}

\section{Attributes \& Skills}

\subsection*{Attributes (1–5)}
\begin{description}
  \item[Body] Strength, endurance, physical force
  \item[Wits] Perception, cleverness, reflexes
  \item[Spirit] Willpower, intuition, resilience
  \item[Presence] Charm, command, social force
\end{description}

\subsection*{Skill Ratings (0–5)}
\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Rating} & \textbf{Meaning} \\
\midrule
0 & Untrained — rely on raw Attribute \\
1 & Familiar — basic competence \\
2 & Skilled — reliable training \\
3 & Expert — professional mastery \\
4 & Master — renowned in your field \\
5 & Legendary — near-mythic talent \\
\bottomrule
\end{tabular}
\end{center}

\section{XP Costs}

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Purchase} & \textbf{Cost} \\
\midrule
Attributes & New rating × 3 XP \\
Skills & New level × 2 XP \\
Followers (On-Screen) & Cap² XP \\
Off-Screen Assets & Minor 4 / Standard 8 / Major 12 XP \\
\bottomrule
\end{tabular}
\end{center}

\section{Deck of Consequences}

Draw from the deck when Complication Points are spent.

\subsection*{Suit Meanings}
\begin{description}
  \item[♡ Hearts] Emotional / Social fallout
  \item[♢ Diamonds] Resource / Wealth loss
  \item[♣ Clubs] Physical harm / Obstacles
  \item[♠ Spades] Mystical / Narrative twists
\end{description}

\subsection*{Rank Severity}
\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Rank} & \textbf{Severity} \\
\midrule
2–5 & Minor setback, scene continues smoothly \\
6–9 & Moderate complication; new obstacle or clock \\
10–K & Severe twist; alters stakes of the scene \\
Ace & Catastrophic turn; reshapes narrative \\
\bottomrule
\end{tabular}
\end{center}

\section{Magic Casting Loop}

\subsection*{1. Channel}
Roll \textbf{Wits + Arcana}:
\begin{itemize}
  \item Each success = 1 Potential
  \item Each 1 = 1 CP immediately
\end{itemize}

\subsection*{2. Weave}
Roll \textbf{Wits + (Art)}:
\begin{itemize}
  \item Apply Description Ladder
  \item Shape spell effect
\end{itemize}

\subsection*{3. Backlash}
GM spends CP for thematic consequences.

\section{Player Archetypes}

\begin{description}
  \item[Solo] Invests in Attributes + Skills. Strong spotlight.
  \item[Mixed] Balances self with one follower or off-screen assets.
  \item[Mastermind] Multiple followers + networks, but more narrative liabilities.
\end{description}

\section{Narrative Time}

\begin{description}
  \item[A Moment] A heartbeat; single action
  \item[Some Time] A few minutes; quick exchanges
  \item[Significant Time] An hour or more; downtime actions
  \item[Days] Extended travel or projects
\end{description}

\section{Boons}

\begin{itemize}
  \item Re-roll one die after seeing the result
  \item Activate an Off-Screen Asset (1 Boon or 2 XP)
  \item Convert 2 Boons = 1 XP (once per session)
  \item Max 5 Boons; overflow converts to XP
\end{itemize}

\section{Travel Framework}

Draw until all suits appear:
\begin{description}
  \item[♠ Spades] Places
  \item[♡ Hearts] Actors
  \item[♣ Clubs] Pressures
  \item[♢ Diamonds] Leverage
\end{description}

Clock Size by highest rank:
\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Rank} & \textbf{Clock Size} \\
\midrule
2–5 & 4 segments \\
6–10 & 6 segments \\
J/Q/K & 8 segments \\
Ace & 10 segments \\
\bottomrule
\end{tabular}
\end{center}

\section{Supply Clock}

\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Segments} & \textbf{Effect} \\
\midrule
0 (Full) & No penalties \\
2 (Low) & Minor complications \\
3 (Dangerous) & Each PC gains Fatigue \\
4 (Empty) & Severe penalties; starvation \\
\bottomrule
\end{tabular}
\end{center}

\section{GM Guidance at a Glance}

\begin{itemize}
  \item Spend Complication Points to add story problems, not punish
  \item Always tie consequences back to thematic domains (♡ ♢ ♣ ♠)
  \item Encourage Intricate actions: reward description with rerolls and narrative control
  \item Let Off-Screen assets resolve downtime problems but keep adventures on the table
\end{itemize}

\section{Asset Activation}

To activate any Off-Screen Asset:
\begin{itemize}
  \item \textbf{Cost:} 1 Boon or 2 XP
  \item \textbf[Scope:] One clear outcome the asset can plausibly produce
  \item \textbf[Limits:] One activation per distinct ask; no auto-wins
\end{itemize}

\section{Follower Assistance}

\begin{itemize}
  \item \textbf{In-Role Assist:} +Cap bonus (plausible fiction)
  \item \textbf{Off-Role Assist:} +1 bonus (intricate description required)
  \item \textbf{Cap:} Maximum +3 total assist bonus per roll
\end{itemize}

\section{Initiative Actions}

Once per scene (crew-wide):
\begin{itemize}
  \item Scout \& Signal — Ally starts Controlled or +1 effect
  \item Distract \& Draw — Reduce kinetic rail by –1 tick
  \item Fetch \& Carry — Move object through danger
\end{itemize}
\textbf{Cost:} Mark Exposure +1 or Harm 1 on follower

\section{Condition Tracks}

\subsection*{Assets/Followers}
\begin{description}
  \item[Maintained] Full capability; no penalties
  \item[Neglected] -1 die when used; requires attention
  \item[Compromised] Unavailable; narrative consequences
\end{description}

\subsection*{Characters}
\begin{description}
  \item[Fatigue (1–4)] — Re-roll successes; 4 = Collapse
  \item[Harm (1–3)] — Narrative injury; 3 = Out of action
\end{description}

\section{Over-Stack Rule}

If crew enters scene with 2+ structural advantages:
\begin{itemize}
  \item Start one named rail at +1, OR
  \item GM banks +1 CP for first ♠ Twist
\end{itemize}

\section{Position and Effect}

\subsection*{Position}
\begin{description}
  \item[Controlled] — Low risk, high chance of success
  \item[Risky] — Moderate risk, balanced stakes
  \item[Desperate] — High risk, likely consequences
\end{description}

\subsection*{Effect}
\begin{description}
  \item[Limited] — Minor impact
  \item[Standard] — Expected result
  \item[Great] — Significant or lasting impact
\end{description}

\section{Remember}

\begin{center}
\textbf{Every roll is a story beat.} \\
\textbf{Embrace the risk—and let the consequences unfold.}
\end{center}
