% !TEX root = ../fates_edge_players_guide.tex

\chapter{Talents}

In \textbf{Fate’s Edge}, \textbf{Talents} are unique abilities that expand what your character can do. They are purchased with XP and often serve as stepping stones toward Prestige Abilities. Talents can be general, cultural, or narrative capstones that reshape how you engage with the world.

\section{Types of Talents}

Talents are organized into three categories based on cost and impact:

\subsection*{Early Talents (3–5 XP)}

Minor mechanical edges or rerolls that enhance niche situations.

\textbf{Examples:}
\begin{itemize}
  \item \textbf{Battle Instincts} — Once per scene, re-roll a failed defense roll.
  \item \textbf{Silver Tongue} — Gain +1 die when persuading or deceiving through speech.
  \item \textbf{Iron Stomach} — Immune to mundane poisons; halve Complications from toxins.
  \item \textbf{Versatile} (Human only) — Once per session, swap one untrained Skill for Trained.
\end{itemize}

\subsection*{Mid-Tier Talents (6–10 XP)}

Significant edges or codified outcomes that deepen specialization.

\textbf{Examples:}
\begin{itemize}
  \item \textbf{Backlash Soothing} (Wood Elf) — Once per session, reduce magical Backlash by 2 CP in natural terrain.
  \item \textbf{Blood Memory} (Ykrul) — After a battle, gain one temporary Skill die reflecting a foe’s tactics.
  \item \textbf{Commanding Presence} — Allies gain +1 to morale rolls; re-roll one failed die when obeying your orders.
  \item \textbf{Familiar Bond} — Gain a magical familiar (counts as On-Screen follower).
\end{itemize}

\subsection*{Prestige Abilities (12+ XP)}

Narrative-capstone powers available only at high investment. These represent cultural myths and pinnacles of mastery.

\textbf{Examples:}
\begin{itemize}
  \item \textbf{Echo-Walker} (High Elf) — Step briefly into Aerisahl; once per arc, turn any Complication into a boon.
  \item \textbf{Warglord} (Ykrul) — Rally scattered warbands into a single host; once per campaign, unify tribes.
  \item \textbf{Spirit-Shield} (Dwarf) — Once per session, erase up to 3 CP from an ally’s roll, taking 1 Backlash.
  \item \textbf{Shadowbinder} (Tulkani) — Once per scene in dim/shadow, negate visibility or tracking entirely.
\end{itemize}

\section{Cultural Talents}

Many Talents are tied to specific cultures or Affinities, reflecting the unique arts and philosophies of a people.

\subsection*{Examples by Culture}

\begin{description}
  \item[Dwarves] — Stone-Sense, Spirit Shield, Forge-Patriarch
  \item[Wood Elves] — Backlash Soothing, Ranger’s Step, Wild Speaker
  \item[High Elves] — Lorekeeper, Weave Anchor, Echo-Walker
  \item[Ykrul] — Cold Fury, Blood Memory, Warglord
  \item[Humans] — Versatile, Guild Ties, Banner-Bearer
\end{description}

\section{Purchasing Talents}

Talents are purchased with XP and may have prerequisites:

\begin{itemize}
  \item \textbf{Attribute/Skill Requirements} — Must meet minimum ratings (permanent, not temporary).
  \item \textbf{Asset Requirements} — Some Talents require ownership of specific assets.
  \item \textbf{Cultural Requirements} — Certain Talents are restricted to specific cultures or Affinities.
\end{itemize}

\textbf{Example:} \textit{Echo-Walker} requires Wits 5, Arcana 4, and High Elf Affinity.

\section{Using Talents}

Talents are typically used once per scene, session, or arc, as noted in their description. Overuse is restricted to prevent imbalance.

\begin{itemize}
  \item \textbf{Once per Scene} — Can be used multiple times per session, but not on the same action.
  \item \textbf{Once per Session} — Limited to one use between downtimes.
  \item \textbf{Once per Arc/Campaign} — Reserved for narrative milestones.
\end{itemize}

\section{Losing Talent Access}

If you later lose a prerequisite (e.g., an Attribute is reduced or an Asset is lost), you \textbf{keep the Talent} but may not be able to use features that require that prerequisite until it is restored.

\section{Summary}

Talents are the unique expressions of your character’s growth:

\begin{itemize}
  \item Early Talents offer small but reliable edges.
  \item Mid-Tier Talents deepen specialization and identity.
  \item Prestige Abilities reshape the story at key moments.
\end{itemize}

Choose Talents that reflect your character’s journey—and let them echo through the world.

