% !TEX root = ../fates_edge_players_guide.tex

\chapter{Asset-Linked Backgrounds}
\label{ch:backgrounds}

In \textbf{Fate's Edge}, your character's past is not just flavor—it's \textbf{function}. \textbf{Asset-Linked Backgrounds} are optional, lore-forward packages that tie your character to a specific place, faction, or tradition. Each requires ownership of a specific \textbf{Asset} and grants a \textbf{Talent}, \textbf{Perk}, and narrative hooks that shape how you engage with the world.

\section{What Is a Background?}
\index{Backgrounds}

A Background is a narrative package that includes:

\begin{itemize}
  \item \textbf{Prerequisite Asset} — You must own or acquire the listed Asset.
  \item \textbf{Talent} — A full-cost Talent tied to the background's theme.
  \item \textbf{Perk} — One mechanical benefit, usable once per session.
  \item \textbf{Scene Keys} — Two venues where this background naturally appears.
  \item \textbf{Hooks} — Narrative pressure points where the GM can spend CP.
  \item \textbf{Invite} (optional) — A behavior cue that may earn a Boon.
\end{itemize}

Backgrounds do not add clocks, stance edits, or extra actions. They are tools for \textbf{roleplay and consequence}.

\section{Example Background: Vhasian Gentry}
\index{Backgrounds!Vhasian Gentry}

\textbf{Requires Asset:} Vhasian Manor, Estate, or County

\textbf{Talent (4 XP):} Influence at Court (Vhasian) — +1 effect on formal petitions and protocol inside Vhasian courts and councils.

\textbf{Perk (Choose 1, once/session):}
\begin{itemize}
  \item +1 effect on one opener in a keyed venue (ally or self).
  \item Hazard −1 for one beat in a named environment.
  \item Mitigate CP for the first on-theme twist.
  \item Restore 1 follower (clear 1 Exposure or heal Harm 1).
  \item Free upkeep on one owned Asset.
\end{itemize}

\textbf{Scene Keys:} Court/Audience, Noble Reception

\textbf{Hooks (GM picks 1 on CP spend in keyed scenes):}
\begin{itemize}
  \item ♣ A rival house "assists" by rerouting your petition to their gatekeeper.
  \item ♢ Protocol demands a costly courtesy now (gift, sponsor, or public apology).
  \item ♡ A jealous clerk bumps another case ahead; you owe someone a face to keep momentum.
\end{itemize}

\textbf{Invite (optional):} "Defer to protocol before acting."

\section{Background Types}
\index{Backgrounds!types}

Backgrounds are grouped by region and theme. Below are examples from various cultures:

\subsection*{Vililan Backgrounds}
\index{Backgrounds!Vililan}

\begin{description}
  \item[Viterra Fenward Bailiff] — Tied to marsh law and abbey rights.
  \item[Ubral Iron-Host Veteran] — Veteran of frontier garrisons and winter drills.
  \item[Silkstrand Span-Factor] — Master of bridges, processions, and quay politics.
  \item[Ecktorian Archivist] — Keeper of stacks, vaults, and forbidden knowledge.
\end{description}

\subsection*{Aeler (Dwarven) Backgrounds}
\index{Backgrounds!Aeler}

\begin{description}
  \item[Aeler Tower Savant] — Scholar of arcane theory and forbidden rites.
  \item[Oath-Pennant Keeper] — Guardian of hospitality and sanctuary law.
  \item[Sepulcher Adept] — Keeper of crypts, bells, and the restless dead.
\end{description}

\subsection*{Linnic and Theona Backgrounds}
\index{Backgrounds!Linnic and Theona}

\begin{description}
  \item[Linnic Tower Scholar] — Master of star-reckoning and numerical harmony.
  \item[Theonan Beacon-Keeper] — Guardian of coastal light and storm warnings.
\end{description}

\subsection*{Kahfagian and Mistlands Backgrounds}
\index{Backgrounds!Kahfagian and Mistlands}

\begin{description}
  \item[Kahfagian Spur Captain] — Master of steppe law and relay routes.
  \item[Mistlands Warden] — Keeper of bells, salt, and the Direwood border.
\end{description}

\subsection*{Cultural and Prestige Backgrounds}
\index{Backgrounds!cultural and prestige}

Some backgrounds are tied to specific cultures or Tier III Assets:

\begin{description}
  \item[Tulkani Pit-Runner] — Champion of quarry trials and masked contests.
  \item[Ykrul Moor Warden] — Guardian of fog, barrows, and old warnings.
  \item[Grand Forge-Master] (Prestige) — Sovereign of heat and hammer.
  \item[Storm-Beacon Lord] (Prestige) — Master of lighthouses and sea-lanes.
\end{description}

\section{Creating Your Own Background}
\index{Backgrounds!creating}

To create a custom Background:

\begin{enumerate}
  \item \textbf{Name It} — Give it an evocative title.
  \item \textbf{Asset Requirement} — Tie it to a Minor/Standard/Major Asset.
  \item \textbf{Talent} — Choose or design a thematically appropriate Talent.
  \item \textbf{Perk} — Pick one mechanical benefit (session-limited).
  \item \textbf{Scene Keys} — Name two venues where it naturally appears.
  \item \textbf{Hooks} — List 2–3 CP sinks with suit markers (♡ ♢ ♣ ♠).
  \item \textbf{Invite} (optional) — Behavior cue that may earn a Boon.
\end{enumerate}

\section{Summary}

Backgrounds in Fate's Edge are about \textbf{belonging}:

\begin{itemize}
  \item They tie you to a place, faction, or tradition.
  \item They offer mechanical benefits and narrative pressure.
  \item They make the world react to your presence.
\end{itemize}

Choose a Background—or craft your own. Let your past shape your future.

\end{chapter}
