% !TEX root = ../fates_edge_players_guide.tex

\chapter{Gods and Powers}
\label{ch:gods-powers}

In \textbf{Fate's Edge}, religion is not just faith—it is \textbf{faction}. The gods and powers of the world are active forces, shaping cultures, inspiring zealots, and demanding tribute. This chapter explores the major divine and mystical forces, their followers, and how they influence the world.

\section{Faith as Faction}
\index{Faith as faction}

Religion in Fate's Edge is deeply political. Each power has its own churches, cults, schisms, and agendas. To serve a god is to join a faction—with all the alliances, enemies, and obligations that come with it.

\section{Major Powers}
\index{Powers!major}

\subsection*{Everflame}
\index{Powers!Everflame}

\textbf{Domains:} Fire, purification, law \\
\textbf{Seat:} Ecktoria \\
\textbf{Themes:} Order, sacrifice, judgment

The Everflame is the dominant faith of the northern shores. It teaches that truth burns away falsehood, and that only through suffering can the soul be purified. Its priests are inquisitors, judges, and executioners.

\textbf{Cleric Talents:} Everflame Keeper, Zealot's Blade, Purifying Light

\subsection*{Light}
\index{Powers!Light}

\textbf{Domains:} Revelation, mercy, community \\
\textbf{Seat:} Viterra \\
\textbf{Themes:} Healing, unity, evangelism

A reformation of the Everflame, the Light emphasizes compassion and community over punishment. Its followers are healers, teachers, and missionaries who seek to unite the faithful under a single banner.

\textbf{Cleric Talents:} Healing Light, Community Blessing, Evangelist's Tongue

\subsection*{Khemesh and Ráeyn}
\index{Powers!Khemesh and Ráeyn}

\textbf{Domains:} Krakens, storms, the sea \\
\textbf{Seat:} Kahfagia \\
\textbf{Themes:} Power, survival, dominion

The twin patrons of Kahfagia, Khemesh and Ráeyn are worshipped by sailors, pirates, and storm-riders. Their temples are built on cliffs and islands, and their priests read the will of the deep in the bones of whales.

\textbf{Cleric Talents:} Storm's Favor, Kraken's Blessing, Deep Oracle

\subsection*{Dhahar, Lord of the Sky and War}
\index{Powers!Dhahar}

\textbf{Domains:} Sky, war, honor \\
\textbf{Seat:} Dhahara \\
\textbf{Themes:} Battle, banners, judgment

Dhahar is the god of the open sky and righteous war. His followers swear by banners and settle disputes with steel. In the south, to break an oath sworn under his name is to invite his wrath.

\textbf{Cleric Talents:} Banner of War, Sky's Judgment, War-Chanter

\subsection*{Ikasha, She Who Sleeps Within Shadow}
\index{Powers!Ikasha}

\textbf{Domains:} Shadow, secrets, transformation \\
\textbf{Seat:} Tulkani underground \\
\textbf{Themes:} Mystery, change, forbidden knowledge

Ikasha is whispered more than preached. Her followers are spies, assassins, and those who walk between worlds. Her temples are hidden, and her rites are known only to the initiated.

\textbf{Cleric Talents:} Shadow Walk, Secret Keeper, Veil of Transformation

\subsection*{The Old Ancestors (Dwarven Faith)}
\index{Powers!Old Ancestors}

\textbf{Domains:} Kin, stone, memory \\
\textbf{Seat:} Aeler mountain halls \\
\textbf{Themes:} Tradition, craft, lineage

The dwarves venerate their ancestors as guardians of the deep laws. To speak an ancestor's name is to invoke their protection—and their expectations.

\textbf{Cleric Talents:} Ancestral Guidance, Stone Communion, Kin-Binding Oath

\section{Other Powers and Spirits}
\index{Powers!other}

\begin{description}
  \item[Ykrul Sky-Spirit] \index{Powers!Ykrul Sky-Spirit} — Nameless shamanic power tied to omens and war.
  \item[Valewood Entities] \index{Powers!Valewood Entities} — Ancient forest powers that shift with the seasons.
  \item[Fae Courts] \index{Powers!Fae Courts} — Elusive beings of whim and consequence.
  \item[The Hunger Below] \index{Powers!Hunger Below} — Primordial force stirring beneath the world.
\end{description}

\section{Divine Service and Consequences}
\index{Divine service}

Serving a god in Fate's Edge is not without cost:

\begin{itemize}
  \item \textbf{Obligations} — Divine mandates, rituals, and tithes.
  \item \textbf{Complications} — Gods may demand sacrifices or intervene in dramatic ways.
  \item \textbf{Blessings} — Divine favor grants Talents, Boons, or narrative edges.
\end{itemize}

\section{Cleric and Divine Spellcasting}
\index{Spellcasting!divine}

Characters who serve the divine may access \textbf{Faith-based Talents} and \textbf{Thaumaturgy}—a form of spellcasting tied to divine will rather than arcane study.

\textbf{Thaumaturgy Arts:}
\begin{itemize}
  \item Purifying Light
  \item Divine Shield
  \item Blessing of Courage
  \item Wrath of the Heavens
\end{itemize}

\section{Summary}

The gods of Fate's Edge, through their worshippers, are active players in the world:

\begin{itemize}
  \item Faith is faction, with politics and power struggles.
  \item Divine service brings both blessings and burdens.
  \item Gods shape culture, law, and the fate of nations.
\end{itemize}

Choose your faith wisely—and be prepared to pay its price.

\end{chapter}
