% !TEX root = ../fates_edge_players_guide.tex

\chapter{Character Advancement}
\label{ch:advancement}

In \textbf{Fate's Edge}, growth is not just about getting better at fighting or sneaking—it's about shaping your character's identity and influence in the world. Advancement comes through \textbf{Experience Points (XP)} \index{Experience Points (XP)}, which you earn through play and spend to enhance your capabilities.

\section{Earning XP}
\index{Experience Points (XP)!earning}

At the end of each session or story arc, the Game Master awards XP based on your actions and choices. Typical sources include:

\begin{itemize}
  \item \textbf{Attendance} — Showing up to the table.
  \item \textbf{Objectives Reached} — Completing major goals or plot points.
  \item \textbf{Discoveries} — Unlocking lore, secrets, or new regions.
  \item \textbf{Hard Choices} — Embracing moral or emotional dilemmas.
  \item \textbf{Complication Spotlight} — Leaning into narrative twists and risks.
  \item \textbf{Bond/Flag Driven Play} — Engaging personal storylines.
  \item \textbf{GM Curveball Award} — For embracing unexpected challenges.
\end{itemize}

\subsection*{Session Pacing Options}

The GM may choose from three pacing modes:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Mode} & \textbf{XP per Session} \\
\midrule
Gritty & 4–6 XP \\
Standard & 6–10 XP \\
Heroic & 10–14 XP \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Session Awards}
\begin{itemize}
    \item Table Attendance: +2 XP
    \item Major Objective Reached: +2–4 XP
    \item Discovery or Lore Unlocked: +1–2 XP
    \item Hard Choice Embraced: +1–2 XP
    \item Complication Spotlight: +1–3 XP
    \item Bond/Flag Driven Play: +1–2 XP
    \item GM Curveball Award: +0–3 XP
\end{itemize}

\subsection*{Milestones}
\ begin{itemize}
    \item +8–12 XP to all players at the conclusion of a major story arc.
    \item +2 XP bonus to one player for a signature moment of the arc.
\end{itemize}

\subsection*{Complication Dividend}
\begin{itemize}
    \item Face Card: +1 XP
    \item Ace: +2 XP
\end{itemize}

\section{Spending XP}
\index{Experience Points (XP)!spending}

XP is the universal currency of growth. You may spend it in three broad categories:

\subsection*{1. Enhance Self}

Invest XP to improve your core capabilities:

\begin{itemize}
  \item \textbf{Attributes}: \index{Attribute!improvement} Cost = New Rating × 3 XP. Downtime = new rating in days.
  \item \textbf{Skills}: \index{Skill!improvement} Cost = New Level × 2 XP. Downtime = new level in days.
\end{itemize}

\textbf{Example:} Raising Melee from 2 to 3 costs 3 × 2 = \textbf{6 XP}.

\subsection*{2. Acquire Assets}
\index{Assets}

XP can be spent to gain worldly influence:

\begin{itemize}
  \item \textbf{Off-Screen Assets} — Minor (4 XP, 1 day), Standard (8 XP, 1 week), Major (12 XP, 1 month).
  \item \textbf{On-Screen Followers} \index{Followers} — Allies who assist in scenes. Cost = Cap² XP. Downtime = 1–3 days to recruit and brief.
\end{itemize}

\textbf{Example:} A Cap 3 Scout costs 3² = \textbf{9 XP}.

\subsection*{3. Learn Talents}
\index{Talents}

Talents are unique abilities that expand what your character can do:

\begin{itemize}
  \item \textbf{General Talents} — Universal benefits like rerolls or bonuses.
  \item \textbf{Racial/Cultural Talents} — Abilities tied to your heritage.
  \item \textbf{Prestige Abilities} \index{Prestige Abilities} — Capstone powers with significant narrative weight.
\end{itemize}

\section{The Three Paths of XP}
\index{Character paths}

Each path reflects a different approach to growth:

\begin{description}
  \item[Solo Path] 70–90\% Enhance Self, 0–10\% Assets, 0–20\% Talents \\
    Focus on personal mastery. Strong in spotlight scenes, but fragile off-scene.

  \item[Mixed Path] 50–65\% Self, 15–25\% Assets, 15–25\% Talents \\
    Balanced growth. Reliable in most scenes, but upkeep and risk apply.

  \item[Mastermind Path] 25–40\% Self, 35–55\% Assets, 20–40\% Talents \\
    Build networks and influence. Powerful in strategy, but vulnerable to fallout.
\end{description}

\section{Downtime and Rushed Growth}
\index{Downtime}

Most advancement requires \textbf{Downtime} to reflect training, study, or negotiation:

\begin{itemize}
  \item \textbf{Attributes}: New rating in days.
  \item \textbf{Skills}: New level in days.
  \item \textbf{Followers/Assets}: 1–30 days depending on scope.
\end{itemize}

\subsection*{Rush Rule}
\index{Rush Rule}

You may skip downtime, but the GM creates a \textbf{Haste Clock} (4 segments). If it fills, the new ability or asset carries flaws or narrative complications.

\section{Tiers of Reputation}
\index{Tiers of Reputation}

As you spend XP, you rise through soft "tiers" that affect how the world responds:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Tier} & \textbf{XP Range} \\
\midrule
I – Rookie & 0–40 XP \\
II – Seasoned & 41–90 XP \\
III – Veteran & 91–150 XP \\
IV – Paragon & 151–220 XP \\
V – Mythic & 221+ XP \\
\bottomrule
\end{tabular}
\end{center}

Higher tiers unlock access to Prestige Abilities and deeper narrative influence.

\section{Follower Upkeep and Risk}
\index{Followers!upkeep}

Followers require maintenance to remain effective:

\begin{itemize}
  \item \textbf{Upkeep}: Each Downtime, pay XP equal to Cap or spend a Scene tending the relationship.
  \item \textbf{Risk}: If the GM spends 2+ Complication Points on an action you take with assistance, they may endanger, injure, or separate the follower instead of you if fictionally appropriate.
  \item \textbf{Off-Screen Capability}: Once per downtime, a Cap 5 follower can solve one significant problem but generates 1 CP for party.
\end{itemize}

\section{Summary}

Advancement in Fate's Edge is about \textbf{choice}:

\begin{itemize}
  \item Will you master yourself?
  \item Will you command networks?
  \item Will you become a mythic figure?
\end{itemize}

Every XP spent reshapes your story. Choose wisely—and live with the consequences.

\end{chapter}
