% !TEX root = ../fates_edge_players_guide.tex

\chapter{Attributes and Skills}

In \textbf{Fate’s Edge}, your character is defined not by a long list of numbers—but by the interplay of four core \textbf{Attributes} and a focused set of \textbf{Skills}. Together, they shape your dice pools, guide your narrative choices, and reflect who you are in the world.

\section{Attributes}

Attributes represent broad aspects of your character’s being. Each is rated from 1 to 5 for most mortals. Exceptional beings may exceed this range.

\begin{description}
  \item[Body] — Strength, endurance, and physical force. \\
    Used for: Melee, Athletics, Endurance.

  \item[Wits] — Perception, cleverness, and reaction speed. \\
    Used for: Arcana, Stealth, Insight, Tactics.

  \item[Spirit] — Willpower, intuition, and resilience. \\
    Used for: Meditation, Resolve, Faith, Endurance.

  \item[Presence] — Charm, command, and social force. \\
    Used for: Diplomacy, Command, Performance, Sway.
\end{description}

\section{Skills}

Skills are focused expressions of talent. Each is tied to one or more Attributes, but creative players may justify alternative pairings if the narrative fits.

\subsection*{Skill Ratings}

\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Rating} & \textbf{Meaning} \\
\midrule
0 & Untrained — Rely on raw Attribute alone \\
1 & Familiar — Basic competence, a journeyman’s touch \\
2 & Skilled — Trained and reliable in most circumstances \\
3 & Expert — Professional mastery or long experience \\
4 & Master — Renowned; your work is widely recognized \\
5 & Legendary — Near-supernatural; few mortals reach this level \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Common Skills}

Below is a list of common skills used in Fate’s Edge:

\begin{description}
  \item[Athletics] — Climbing, running, swimming.
  \item[Arcana] — Magical theory, rituals, spellwork.
  \item[Brawl] — Fists, grappling, improvised fighting.
  \item[Insight] — Intuition, empathy, lie detection.
  \item[Melee] — Blades, axes, polearms.
  \item[Ranged] — Bows, crossbows, thrown arms.
  \item[Diplomacy] — Negotiation, mediation, etiquette.
  \item[Stealth] — Hiding, shadowing, evading.
  \item[Deception] — Disguise, misdirection, bluffing.
  \item[Survival] — Tracking, foraging, navigation.
  \item[Command] — Leadership, intimidation, rallying.
  \item[Craft] — Smithing, alchemy, tinkering.
  \item[Performance] — Music, oratory, storytelling.
  \item[Lore] — History, cultures, languages.
\end{description}

\section{Building Your Dice Pool}

When attempting a significant action, combine the relevant Attribute + Skill to determine your dice pool.

\textbf{Examples:}
\begin{itemize}
  \item Body + Melee = raw steel and sinew.
  \item Wits + Melee = anticipation and precision.
  \item Presence + Melee = flourish, feint, and intimidation.
\end{itemize}

The same Skill, seen through different Attributes, tells a different story.

\section{Skill Caps and Growth}

Each Skill is capped by its linked Attribute. You cannot raise a Skill above your Attribute rating.

\textbf{Cost to Raise Skills:} New level × 2 XP

\textbf{Example:} Raising Stealth from 2 to 3 costs 3 × 2 = \textbf{6 XP}.

\section{Skill Specializations}

Some skills may be used in specialized contexts. Players may describe niche expertise (e.g., “Desert Survival” or “City Stealth”) to gain narrative advantages, subject to GM approval.

\section{Summary}

Attributes and Skills are the core of your character’s identity:

\begin{itemize}
  \item Attributes define your potential.
  \item Skills reflect your training and focus.
  \item Together, they determine how you act—and how the world responds.
\end{itemize}

Choose wisely. Your capabilities are not just tools—they are expressions of who you are.

