% !TEX root = ../fates_edge_players_guide.tex

\chapter{World Regions}
\label{ch:world-regions}

The world of \textbf{Fate's Edge} is vast, diverse, and shaped by ancient empires, clashing cultures, and the ever-present threat of the unknown. This chapter explores the major regions around the \textbf{Amaranthine Sea}—the heart of civilization—and beyond.

\section{The Amaranthine Sea}
\index{Amaranthine Sea}

At the center of the western world lies the \textbf{Amaranthine Sea}, a vast inland body of water that has been the cradle of empires, the highway of trade, and the battleground of kings. Its shores are home to a mosaic of cultures, each with its own traditions, ambitions, and dangers.

\section{Northern Shores: Vililan}
\index{Vililan}

The western and northern coasts of the Amaranthine are dominated by \textbf{Vililan}, the remnants of the once-great Utaran Empire. Here, old roads still bind the markets, and ancient laws still hold sway—even if their enforcers are long gone.

\subsection*{Major Regions}

\begin{description}
  \item[Ecktoria] \index{Ecktoria} — The rump of the old empire, clinging to marble and ceremony. Home to gladiatorial arenas and the Church of the Everflame.
  \item[Acasia] \index{Acasia} — A fractured province of warlords and bravo companies. Silkstrand, its great port, is a beacon of cosmopolitan vice and trade.
  \item[Vhasia] \index{Vhasia} — A land of dukes, bloodlines, and old rivalries. Honor and oaths still matter—though not always as they should.
  \item[Viterra] \index{Viterra} — Highland pragmatists and fen traders. Known for their Knights of the Dawn and a culture of logistics.
  \item[Ubral] \index{Ubral} — Highland clans and dwarven allies. Guerrilla fighters who know every sheep-track and scree path.
  \item[Thepyrgos] \index{Thepyrgos} — The last lantern of learning, where orators and scholars debate the fate of the world.
  \item[Mistlands] \index{Mistlands} — A fog-shrouded protectorate where dwarves and humans coexist uneasily with stranger things.
  \item[Kahfagia] \index{Kahfagia} — A maritime oligarchy that commands the Titan's Throat and the Dolmis Sea. Storms and krakens are their allies.
\end{description}

\section{Southern Reaches: Akilan}
\index{Akilan}

To the south lies \textbf{Akilan}, a continent of sun, stone, and river. Here, monsoon-fed empires rise and fall, and the power of scribes and priests rivals that of kings.

\subsection*{Notable Nations}

\begin{description}
  \item[Oshiira] \index{Oshiira} — A bureaucratic machine of canals and granaries. Their prefects and census-takers hold more power than generals.
  \item[Sekogo] \index{Sekogo} — Druidic jungle cities and lagoon ports. Masters of river and root.
  \item[Taharka] \index{Taharka} — Terrace highlands and hydraulic works. Divine mandate guides their rulers.
  \item[Ashaan] \index{Ashaan} — Slaver-empire turned fractured successor states. Gold, ivory, and ancient tombs draw adventurers and fools alike.
\end{description}

\section{Eastern Horizons: Ostrilan}
\index{Ostrilan}

Across the sea to the east lies \textbf{Ostrilan}, a vast land of monsoons, old empires, and steppe riders. Here, the winds carry the scent of incense and the clash of distant wars.

\subsection*{Key Regions}

\begin{description}
  \item[Dhahara] \index{Dhahara} — Peninsula of sky and memory. Monsoon ports and desert kingdoms vie under the banner of Dhahar, Lord of War.
  \item[Sihai] \index{Sihai} — A thousand terraces and ten thousand seals. River dynasties and scholar-magistrates shape this vast empire.
  \item[Nihon] \index{Nihon} — Isles of quiet steel. Mountain domains and sea-lords uphold the way of the blade.
  \item[Ayohkhan] \index{Ayohkhan} — Emerald archipelago of smoke and spice. Jungle thrones and monsoon fleets rule the tides.
\end{description}

\section{The Wilds Beyond}
\index{Wilds}

Beyond the coasts lie lands less mapped and more dangerous:

\begin{description}
  \item[Valewood] \index{Valewood} — Immense, uncanny forest where geography shifts and old powers linger.
  \item[Ykrul Steppes] \index{Ykrul Steppes} — Vast grasslands where nomad warbands ride under open sky.
  \item[The Crimson Basin] \index{Crimson Basin} — Rainforest heart of Akilan, contested between Wood Elves and settlers.
  \item[The Great Desert] \index{Great Desert} — Harsh and unforgiving, dotted with oases and lost kingdoms.
\end{description}

\section{Travel and Trade Routes}
\index{Travel routes}

\begin{itemize}
  \item \textbf{The Titan's Throat} \index{Titan's Throat} — The narrow strait that connects the Amaranthine to the outer ocean. Controlled by Kahfagia.
  \item \textbf{The Belworth Corridor} \index{Belworth Corridor} — Main artery between Vhasia and Viterra.
  \item \textbf{The Way of Silk} \index{Way of Silk} — Southern trade routes linking Fhara, Kuvani, and Dhahara to northern ports.
  \item \textbf{The Khesai River} \index{Khesai River} — Spine of Ashaan trade and power.
\end{itemize}

\section{Summary}

The world of Fate's Edge is vast and varied:

\begin{itemize}
  \item Each region offers unique cultures, conflicts, and opportunities.
  \item Trade routes connect distant lands—and also spread danger.
  \item The past is never truly gone; it waits in ruins, tombs, and old oaths.
\end{itemize}

Where will your story begin? And what will you leave behind?

\end{chapter}
