% !TEX root = ../fates_edge_players_guide.tex

\chapter{Compendium of Talents}

In \textbf{Fate’s Edge}, \textbf{Talents} are the unique abilities that define your character’s exceptional capabilities. This chapter provides a comprehensive list of Talents organized by tier and theme, from minor edges to mythic powers.

\section{Design Principles}

Talents in Fate’s Edge are designed to:

\begin{itemize}
  \item Enhance narrative positioning rather than raw power.
  \item Create story hooks and complications.
  \item Reflect cultural identity and personal growth.
\end{itemize}

Talents do \textbf{not}:
\begin{itemize}
  \item Pierce the global +3 Assist cap.
  \item Create Diamonds (narrative leverage) directly.
  \item Add extra actions beyond normal Initiative rules.
\end{itemize}

\section{Early Talents (3–5 XP)}

Minor mechanical edges or rerolls that enhance niche situations.

\begin{description}
  \item[Battle Instincts (4 XP)] — Once/scene, re-roll a failed defense check.
  \item[Silver Tongue (3 XP)] — +1 effect on negotiation or fast-talk openers in keyed venues.
  \item[Stone-Sense (5 XP) [Dwarf/Stone-trained]] — Read flaws in stone; once/scene grant start Controlled to a move/breach.
  \item[Iron Stomach (3 XP)] — Resist mundane toxins; Mitigate first CP 1 from poison/rot this scene.
  \item[Versatile (3 XP) [Human]] — Once/session, treat adjacent untrained task as Trained.
  \item[Backlash Soothing (5 XP) [Woodwise]] — Once/scene in natural terrain, reduce magical Backlash by 1 CP.
  \item[Blood Memory (4 XP) [Ykrul]] — After violent scene, name tactic; next counter vs same style gains +1 effect.
  \item[Hand Signals (2 XP)] — When you set silent plan, next ally gains +1 effect.
  \item[Route Whisper (2 XP)] — On declared route, team's first crossing gains +1 effect or Escape −1.
\end{description}

\section{Mid-Tier Talents (6–10 XP)}

Significant edges or codified outcomes that deepen specialization.

\begin{description}
  \item[Commanding Presence (8 XP)] — When ally follows explicit order, start Controlled (once/ally/scene).
  \item[Warden of Lanterns (8 XP)] — Night scenes near you reduce darkness penalties; once/scene name ally's opener Controlled.
  \item[Storm-Caller (8 XP) [Dhaharan/Sihai]] — Once/scene, influence weather: Hazard −1 (wind/rain) or Mitigate CP 1.
  \item[Beast-Tongue (8 XP) [Ykrul]] — Parley with local beasts; once/scene treat creature as Cap 2 follower.
  \item[Rune-Keeper (8 XP) [Aeler/Dwarf]] — Once/scene scribe ward: Mitigate first CP 1 or Hazard −1 vs intrusion.
  \item[Blood-Ties Oath (8 XP) [Fhara]] — When oath/honor at stake, first action gains +1 effect and Protected Turn.
  \item[Guild Ledgercraft (8 XP) [Oshiiran]] — Once/scene at market, start Controlled for settlement or Mitigate CP 1.
  \item[Echo-Courtier (8 XP) [Aeler]] — Once/scene in formal society, shift social rail −1 or Mitigate CP 1 from "memory".
  \item[Battle Chant (8 XP) [Ubral/Viterra]] — Allies ignore first Fatigue 1 from exertion this combat.
  \item[Whisper-Lord (8 XP) [Sekogo/Oshiiran]] — Once/session, ask pointed question about target's status; GM answers truthfully.
  \item[River-Sovereign (8 XP) [Dhaharan]] — Once/scene on water, call right-of-way: Hazard −1 or start Controlled.
  \item[Tide-Walker (8 XP)] — First surge in water scene counts as Controlled; once/scene Hazard −1 on 10-ft square.
\end{description}

\section{Leadership Talents}

Talents that enhance coordination and tactical control.

\begin{description}
  \item[Coordinated Assault (8 XP)] — When you Lead, direct up to two helpers (total Assist still +3).
  \item[Master Coordinator (15 XP)] — Once/scene after successful Lead: +1 effect or open second Initiative window (take Fatigue 1).
  \item[Banner-Bearer (10 XP)] — Once/battle, Mitigate first 2 CP from group result (distributed as narrated).
\end{description}

\section{Prestige Talents (12+ XP)}

Narrative-capstone powers that reshape the story.

\begin{description}
  \item[Echo-Walker (18–20 XP) [Aeler]] — Once/arc, transmute 1–2 CP into Omen: bank +1 effect next action.
  \item[Warglord (18 XP) [Ykrul]] — Once/campaign, unify bands under one banner: treat allies as Cap 3 unit.
  \item[Spirit-Shield (15 XP) [Dwarf]] — Once/scene, reduce ally's CP by up to 2; you mark Fatigue 1 and Backlash 1.
  \item[Shadowbinder (16 XP) [Tulkani/Veil]] — Once/scene in dim/shadow, treat first detection as Controlled and Hunt −1.
  \item[Forge-Patriarch (18 XP) [Aeler/Dwarf]] — Within Great Forge-Citadel, set cadence: one craft sequence starts Controlled.
  \item[Master of Coin (16 XP)] — When presiding at exchange, settlements gain +1 effect; once/session Mitigate CP 1.
  \item[Warden of Storms (18 XP) [Sihai/Dhaharan]] — Once/arc, anchor/dispose major storm; your side's first two Hazards −1.
  \item[Mask-Bearer (16–17 XP) [Nihon/Masked Courts]] — Once/scene while masked, escape recognition or gain Protected Turn.
  \item[Star-Reader (18 XP) [Linnic/Sihai]] — Once/arc under predicted sky, one action gains start Controlled and +1 effect.
  \item[Everflame Keeper (16 XP) [Ecktoria]] — Once/scene, sanctify/purify: treat corruption Hazard as Hazard −1 + Protected Turn.
  \item[Resonant Architect (18 XP) [Dwarf]] — Once/arc, animate place's will: treat venue as Cap 3 follower for one scene.
  \item[Sandseer (16–18 XP) [Fhara]] — Once/scene in deserts, cancel/cvt enviro Hazard tick to +1 effect.
\end{description}

\section{Magic Specializations}

Talents that enhance spellcasting and magical arts.

\begin{description}
  \item[Elemental Focus (8 XP)] — Pick element (Fire/Water/Earth/Air). Once/scene: +1 effect or Mitigate Backlash 1.
  \item[Backlash Resistant (8–10 XP)] — First Backlash each scene reduced by 1 CP (min 0).
  \item[Ritual Master (12 XP)] — Lead multi-participant rituals without escalating Backlash tier; once/ritual convert 1 CP to Fatigue 1.
  \item[Spell Anchor (8 XP)] — Once/session, stabilize failing spell by taking Fatigue 1; spell outcome stays current.
  \item[Conduit Walker (15 XP)] — Once/arc, open conduit to Otherwhere: bounded question or subtle boon.
\end{description}

\section{Cultural Skill Paths}

Each culture has associated skill focuses:

\begin{description}
  \item[Scholarly Arts] — Lore, Research, Arcana, Craft, Ritual
  \item[Social Arts] — Diplomacy, Sway, Performance, Subterfuge, Insight
  \item[Martial Arts] — Melee, Ranged, Brawl, Athletics, Tactics
  \item[Wilderness Arts] — Stealth, Survival, Navigation, Animal Handling, Scouting
\end{description}

\section{Summary}

Talents are the unique expressions of your character's journey:

\begin{itemize}
  \item Early Talents offer small but reliable edges.
  \item Mid-Tier Talents deepen specialization and identity.
  \item Prestige Talents reshape the story at key moments.
\end{itemize}

Choose Talents that reflect your character's growth—and let them echo through the world.

