\chapter{The Vent That Whispered}

\section*{Told by}

Old Orra, lamp−keeper of the Third Stack, Aelerian Range. Not been underground in thirty years, but still counts her breaths before sleep. She told this tale in the dark. Would not light a candle.

\section*{The Tale}

You ever walk the Deep Galleries alone? No. Good. Don't start.

The air down there has a measure. Eight breaths is a corridor. Sixteen is a chamber. Thirty−two is a fool who forgot to turn back. That's what they taught me. Breathe, count, turn. Never answer the ninth.

I was young. Lamp trimmed, chalk dry. Third Stack to the old sump. Easy walk. Should've been two hours.

At the junction −−− you know the one, three tunnels, the middle one always drips −−− I stopped. Something was wrong. The drip had stopped. The echo had changed. The stone felt soft under my boot. And then I heard it.

Not with my ears. Inside. Like a memory that wasn't mine. Like someone else's grief pressed into my skull like a thumb into clay.

\emph{``Count the ninth breath.''}

I didn't. I walked faster. The voice followed.

\emph{``The ninth holds the name of the miner who died here. The year the pillar cracked. The route to the lost sump. The name of the man who forgot to turn back.''}

My lamp flickered. The flame burned blue for a heartbeat. I didn't stop to trim it. I kept counting. Eight in. Eight out. My lips moved. The voice couldn't find the crack.

When I reached the surface, the lamp was dry. My shadow was gone. The sun was setting and I cast nothing. The other miners stared. They did not ask.

The next morning I looked in a mirror −−− a real one, polished brass, not a still puddle −−− and the reflection smiled first.

Took a week for the shadow to come back. Took another for my left hand to stop writing in a script I couldn't read. The Vent−Prior said the deep had spoken and I hadn't answered, and that was mercy. She said the mountain respects the ones who know where the door is.

She promoted me to lamp−keeper. I never went down again. I tend the wicks. I count the hours. I do not breathe in rhythm.

\section*{The Witch's Closing}

``Stone remembers. So does the space between breaths. Answer neither. And if you must walk the deep, count to eight −−− then start over. The ninth is not for you.''

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Voice in the Crack}

\paragraph{Description:}

A psychic remnant of a Deep Drake sealed in the mountain. It cannot move, but it whispers into the minds of those who walk alone in deep, dark, air−starved places. It offers secrets like a miser offers silver. It asks for the ninth breath −−− the one you were not meant to take.

\paragraph{Tags / Mechanics:}

\begin{itemize}
ℹtem Tags: [PSIONIC] [DOMAIN (stone)] [DREAD]
ℹtem Harm: none directly −−− mental strain.
ℹtem Fatigue: On failed Spirit + Resolve (DV 4), mark 1 Fatigue and hear a false memory so vivid it feels true for a heartbeat.
ℹtem Timer : \textit{Counting the Ninth} [4] −−− ticks each time a character answers a question while alone in the dark. When full, the character's shadow detaches for 1d3 days. During this time, social Position worsens by one step (people feel something missing), and the shadow may act independently −−− watching, pointing, once opening a door the character meant to keep shut.
ℹtem Strings: The Voice can offer a \textit{True Shem} −−− a forgotten fact, a lost name, a route through stone that the maps erased −−− in exchange for counting the ninth breath (advancing the timer by one segment). The secret is always true. The price is always the same.
\end{itemize}

\paragraph{Motivation:}

Feeds on attention and isolation. Wants to be witnessed. Does not hate the living. Does not love them either. It is lonely. That is worse.

\paragraph{Adventure Hook:}

A vent−warden has gone missing in the Third Stack. The party finds a crack in the wall −−− and hears breathing on the other side. Slow. Measured. Counting.

\section*{Colophon}

Recorded in the Third Stack hostel, winter of 862, by Dusana of the Raven Road. The chalk circle was erased before I slept. My shadow arrived a heartbeat late. I did not ask where it had been.

\noindent \textit{The ink is soot and melted wax. The lamp is borrowed.}