\chapter{The Hollow Bride}

\section*{Told by}

Mabs the Miller, a round woman with flour on her apron and a cider−cup always half full. Tells stories while she works the grindstone. Says the Hollow talks to her at night. Would not say what it says.

\section*{The Tale}

You know about the ninth cup, don't you? Course you do. Every child knows: you pour eight. You never pour the ninth. The ninth is for the Hollow.

But children forget. Or they get curious. Or they think the rules are for other people, the ones who didn't grow up in a house with a red door and a mother who counted every breath. That's how it always starts.

Little Kesta −−− she was what, seven? eight? −−− went to the well at dusk. The adults were at the harvest feast, the kind where the ale flows and the children run wild and no one watches the threshold. She was supposed to draw water for the morning bread. Simple work. She'd done it a hundred times.

She counted the buckets. One, two, three, four, five, six, seven, eight.

She pulled them up. She set them in a row. The water was clear. The moon was not yet up. The well was just a well.

Then she thought: what's one more? Just to see.

She drew the ninth bucket. The rope was cold. The water was black. Not dark −−− black. Like someone had poured ink into the deep and stirred it with a bone. And something climbed out.

It wore her mother's wedding dress, the one that hung in the cedar chest, the one no one had touched since the plague year. It smiled with her mother's mouth, the way her mother smiled when she was tired and pretending not to be. But its eyes were the colour of a drowned moon. Grey. Swollen. Watching.

``Hollow Bride,'' Kesta whispered. That's what the old folk call it. You don't name it. You don't speak its name. You don't give it that kind of hold. She named it.

The Bride followed her home. Walked three paces behind, silent as a held breath. Stood outside the red door. Knocked.

Three knocks. Then a pause. Then three more. The rhythm of a heart that stopped. The rhythm of someone counting to eight and then waiting.

Kesta's mother opened the door. She saw the dress. She saw the mouth. She did not see the eyes −−− she looked at the threshold, not the face. ``Come in,'' she said. ``You must be cold.'' That's the law of hospitality. A red door means bread, salt, and one safe night. You cannot refuse. The door remembers.

The Hollow Bride stepped inside. It sat at the table. It didn't eat. It didn't drink. It didn't speak. It just watched. Watched Kesta braid her hair. Watched the mother bank the fire. Watched the father pretend not to see.

At midnight, it spoke. Its voice was the sound of water dripping in a well that had gone dry.

``You poured the ninth cup. You named me. Now I am your guest.''

The mother understood then. She cut a loaf with a knife that had been blessed at the Moot Oak, the kind of knife that costs three silver and a whispered prayer. She put salt on the table with her left hand, palm up, the way the hedge−witch taught her. She said: ``Guest−right lasts one night. At dawn, you leave.''

The Hollow Bride smiled. It was the mother's smile. It was wrong.

``I will leave,'' it said. ``But I will return. The ninth cup is an invitation. The threshold remembers. The red door remembers. And I have a long memory for small slights.''

Kesta is an old woman now. She still sets eight cups on her table. She still leaves the ninth empty. She still checks the red door at dusk and again at dawn. And every harvest, when the lanterns burn blue at the ford and the sheep will not cross the bridge, she walks the perimeter and touches every stone.

So far, it's held. So far, the Bride has not returned.

But the miller says the well water tasted of salt last spring. And the sheep won't drink.

\section*{The Witch's Closing}

``Hospitality is a thread. Tie it right, and it binds the guest. Tie it wrong, and it binds you. The thread does not fray on its own. Someone cuts it. And never, ever pour the ninth. Not even to see what happens. Because something always does.''

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Hollow Bride}

\paragraph{Description:}

A threshold−spirit that wears the face of a loved one. It can only enter a home if invited −−− either by name, by pouring a ninth cup, by drawing a ninth bucket, or by breaking any taboo that counts to eight and then pauses. It does not hate the living. It is hungry for something else. Recognition, perhaps. Or the sound of a door opening when it should stay shut.

\paragraph{Tags / Mechanics:}

\begin{itemize}
ℹtem Tags: [FAE] [THRESHOLD] [DREAD]
ℹtem Cannot cross a red door unless invited. Repelled by bread, salt, and a knife blessed at a moot or crossroads.
ℹtem Harm: none directly. It imposes Conditions. \textit{Watched} (someone is looking at you from the corner of every room). \textit{Invited} (you spoke its name or poured the ninth −−− it will remember). \textit{Named} (you gave it a name it did not ask for. Names have weight).
ℹtem Timer : \textit{The Ninth Cup} [6] −−− ticks each time hospitality is broken: a door left unlatched, a guest refused bread, a ninth cup poured. When full, the Bride can enter any home in the village. Does not need to knock first.
ℹtem Bargaining: The Bride may be offered a substitute −−− a memory, a name, a firstborn child, the silence of a particular witness −−− to buy another year. The substitute must be offered freely. The Bride does not take. It accepts.
\end{itemize}

\paragraph{Motivation:}

To be invited. To be remembered. To stand outside a red door and hear it open from the inside. It is not cruel. It is patient. That is worse.

\paragraph{Adventure Hook:}

A village has the same nightmare: a woman in a wedding dress outside a red door, knocking. Three knocks. Then a pause. Then three more. The sheep are dying. The well tastes of salt. The party must find who broke guest−right and heal the threshold before the ninth cup is poured again.

\section*{Colophon}

Recorded in the miller's kitchen, year of the sour harvest. The kettle was warm. The cider was dry. The Bride did not knock that night. The miller put an extra log on the fire anyway. Said she could feel something standing in the yard, watching the window. I did not look.

\noindent \textit{The ink is lampblack and rain. The ninth cup remains on the shelf.}