\chapter{The Warrior Who Refused to Call the Runner}

\section*{Told by}

Temur Many-Scars, steppe rider with a braid down to his belt and a laugh like breaking bones. Been in more camps than he can count. Says the only thing that kept him alive was bad geometry. He taught me the kon'reh that night. I lost every match.

\section*{A Note on the Telling}

The Ykrul speak of the board with their own tongue. I have written the Ykrul words in \textit{italics} where they first appear, then given the common term. The kon'reh is forced into the telling because the Ykrul do not separate war from the game that teaches war. You cannot understand one without the other. Temur made this clear. He was polite about it. Politeness on the steppe means you are being taught something. Rudeness means you are being measured.

\medskip
\noindent\begin{tabular}{ll}
\textit{War-Chief} & the piece others call Blue \\
\textit{Runner} & Green \\
\textit{Well} & Sanctum \\
\textit{Calling a Runner} & Seeding \\
\textit{Nailed} & Rooted \\
\textit{The Five-Breath Race} & Reforge \\
\textit{Cuts} & specials \\
\textit{Breaths} & plies \\
\textit{The shadow of the stone} & Zone of Control (ZoC)
\end{tabular}

\section*{The Tale}

The Khagan called a muster. We were young, hungry, and stupid. The kind of stupid that gets carved into the clan archive as a lesson for grandchildren. My cousin --- bold as a winter wolf, twice as fast, half as patient --- wanted glory. He pushed his \textit{War-Chief} forward across the board. He didn't count exits. He just pushed.

The enemy stepped aside. Let him pass. Let him think he was winning. That's the oldest trick on the steppe. Give a fool enough rope and he'll hang himself. Give a warrior enough board and he'll forget the Well.

By the time he reached the \textit{Well}, his \textit{cut} was spent. He looked at the far \textit{Well}. It was empty. He could \textit{Call a Runner}. But his \textit{War-Chief} would be \textit{Nailed} for a \textit{breath}. Rooted. Fixed. A target the size of a tent flap.

I said: ``Don't. Their \textit{War-Chief} still has a \textit{cut}. They're two \textit{breaths} away from boxing you in. Count the shadow before you move.''

He laughed. ``They won't risk it. They're cowards. Look at them. They haven't moved in three beats.''

He \textit{Called}.

The enemy hopped a stone --- just one, the smallest move on the board, the kind you learn on your first day and forget by your second --- landed on the \textit{Well}, and \textit{displaced} his \textit{War-Chief} off the board. The \textit{Five-Breath Race} started ticking. Five beats. That's all my cousin had. Plant his banner in five \textit{breaths} or lose the match and the tribute and the clan's good opinion.

He didn't. He couldn't. The enemy had already welded the approach. Every lane ended in \textit{the shadow of the stone}. The race was dead before it started. The banner never touched ground.

We lost. The clan paid tribute for a year. Silver. Horses. Three strings of good beads that my grandmother had woven. The Khagan made us watch as she handed them over.

The Khagan demoted him to herd-watcher. She took his braid. She took his seat at the fire. She said: ``You forgot the first number. Not the second, not the third. The first. Count exits before you count glory. Call only when the geometry is yours. The board does not forgive. Neither do I.''

My cousin watched the sheep for three winters. He learned to count. He never played kon'reh again.

\section*{The Witch's Closing}

``A free \textit{War-Chief} is a noose waiting for a neck. \textit{Call} only when the shadow is yours. Always keep one \textit{cut} for the door. The board is patient. It will wait for you to forget.''

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Herd-Watcher's Regret (a curse)}

\paragraph{Description:}

A punishment that falls on warriors who \textit{Call} unsafely --- who move without counting exits, who trust luck over geometry. The curse does not strike the body. It strikes the name. Reputation frays. Rank dissolves. Direction fails at the worst moment. The victim becomes forgettable. Still alive. Still breathing. Still eating and sleeping and walking the camp. But no longer counted.

\paragraph{Tags / Mechanics:}

\begin{itemize}
\item Tags: [CURSE] [SOCIAL] [DREAD]
\item Effect: -1 Position on all Command and Sway rolls. People listen but do not hear.
\item Once per session, the GM may force a re-roll on any Navigation or Tactics check. The character remembers the wrong path. The short cut leads to a ravine. The ambush point is already occupied.
\item Strings: The curse can be lifted by a public act of heroic geometry --- winning a kon'reh match against a worthy opponent, or completing a Five-Breath Race under the old rules. No one can help. The geometry must be yours.
\item Timer : \textit{Forgotten Name} [6] --- ticks each time the character gives an order that is ignored, or each time someone looks past them to address someone else. When full, the character loses their name entirely. No one remembers them. Not their mother. Not their khagan. Not the enemy they once defeated. They become a stranger at their own fire.
\end{itemize}

\paragraph{Motivation:}

The curse is not sentient. It is a habit of the steppe, a pattern carved into the bone of the Ykrul by ten thousand kon'reh matches. The board remembers what the players forget. It enforces its own geometry.

\paragraph{Adventure Hook:}

A decorated warrior has disappeared from the camp archive. Not physically. She still sits by the fire. She still eats her share of the meat. But everyone looks past her. The khagan's scribe cannot recall her name. The clan's children do not know her face. The party must restore her name --- or watch her fade into the shadow of the stone.

\section*{Colophon}

Winter camp of the Grey Fox clan, somewhere between the salt flats and the frozen river. The rider spat tea leaves into the fire and said: ``That was the winter I learned to count. The sheep were better company than the board.'' He taught me three openings and one closing. I remember the closing.

\noindent \textit{The ink is tea and ash. The board is still on the table.}