\chapter{The Silk Witness}

\section*{Told by}

Maer of the Root-Circle, a graying wood elf who mends bridges and remembers every broken oath. Speaks like water over stones --- soft, but sharp. Her hands never stop moving. Mending, knotting, untangling.

\section*{The Tale}

A merchant came to the Valewood, full of himself and his archives. He needed a bridge. He had coin. He did not have courtesy.

He spoke to the Lethai-al wardens. "Name your price," he said. "Double if you build it by autumn."

They sent for the Silk Vigil. The Vigil hung a thread across the river. White silk, thin as a spider's. "Cross this," they said, "and we will talk."

The merchant laughed and stepped over.

The thread broke.

The Vigil nodded. "You have refused witness. Build your bridge. Every plank will remember your rudeness. No one will sing its name. No child will skip across it. No lover will meet at its center."

He built it. It held. Stone and good timber. No curse in the mortar. But the workers called it "the Rudeness." Travelers spat on its railings. Mothers pulled their children to the other bank.

Years later, he returned, bent and grey. The bridge had not fallen. It had not forgiven either. He knelt at the Vigil's threshold. "What must I pay?"

The Vigil picked up a single strand of silk from the dust. Held it to the light. "Say this: 'I am sorry for the courtesy I broke.' Say it once, and mean it. Then build a garden under the bridge, for the eels to hide in. Soft places. Mud. Roots. Let the water slow."

He did. He knelt in the mud himself, planting rushes. The eels came. The bridge softened. The moss grew over the stone where men had spat.

Now they sing of it --- not as a curse, but as a repair. Travelers leave a thread tied to the rail. One strand. White. Thin.

\section*{The Witch's Closing}

``Mercy with memory. That is the thread. Cut it, and you build alone. Heal it, and the road belongs to everyone. The eels remember both.''

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Unwitnessed Bridge}

A bridge built without proper rites --- no thread, no witness, no offering of courtesy. Sound as stone, sound as timber. But it radiates a low social curse, thin as a spider's silk, easy to ignore until you feel it in your teeth.

Those who cross feel irritation, suspicion, the sense that someone spoke behind their back and stopped when they turned.

\paragraph{Tags / Mechanics:}

\begin{itemize}
\item Tags: [CURSE] [LOCATION] [SOCIAL]
\item --1 Position on social rolls while on or near the bridge.
\item Once per day, the GM may force a reroll on any Command or Sway roll attempted while crossing. Take the lower result.
\item Clock: \textit{Repair the Bridge} [6] --- ticks when someone crosses without courtesy, or when the bridge is used for argument or threat. Needs an offering of courtesy (bread, salt, a spoken apology) and a public act of repair --- planting, cleaning, mending a cracked rail. Each full segment cleared requires one such act, witnessed by at least three people.
\end{itemize}

\paragraph{Adventure Hook:}

A new bridge divides a river community. Merchants whisper of bad luck. Bargains struck on the bridge sour within a week. Lovers quarrel at the midpoint and do not know why. The builder died rich, unrepentant, and left no heir. The bridge needs healing. The question is who will kneel in the mud.

\section*{Colophon}

Sat under a roof-tree at the edge of the Root-Circle. The eels were listening. I left a thread.

\noindent \textit{The ink is river water and crushed rush. The candle is borrowed.}