\chapter{The Gloss That Ate a Name}

\section*{Told by}

Archivist Velynne, a Lethai with ink under her nails and a voice that never rises above a murmur. Says she has forgotten more languages than most people learn. Her ears are pierced with copper rings. She would not tell me what they commemorated.

\section*{The Tale}

In the great library of Thepyrgos −−− the Lethai do not call it a library, they call it the Still−Room, where words go to be forgotten properly −−− there is a scroll that rewrites itself. Copied from a copy from a copy, each margin a correction, each correction a new error. The Lethai keep it because they keep everything. Even hunger.

A scholar −−− call her Miralis, though that was not her first name and not her last −−− wanted the original. The Still−Room keepers refused. "The original is not lost," they said. "It is eating. We do not feed it."

Miralis found the forbidden room anyway. Lethai doors do not lock. They wait. Inside, a single scroll sealed with wax the color of old bruises. She broke the seal.

The scroll was blank. She turned it over. Blank. She held it to the lamp. Blank.

Weeks passed. Miralis began to forget the taste of rain. Her students' faces blurred into the same pale oval. Her mother's voice became a hum, then a frequency, then nothing. She returned to the Still−Room. The scroll now held one word, written in her own hand though she had never touched a brush:

\textit{Miralis}.

She understood then, the way the Lethai understand hunger: slowly, then all at once. The scroll did not store knowledge. It \emph{ate} it. The copies were not safety. They were rations. The original was the mouth.

She burned it with a lamp she had carried since childhood −−− a brass thing, dented, warm. The fire was cold. The smoke smelled of a birthday she could no longer describe. Her name came back in pieces. Her grandmother's face stayed gone. That was the toll the Still−Room keepers had tried to warn her about.

They did not say \emph{I told you so}. Lethai do not. They poured her tea and watched her drink. The tea had no taste. She could not remember if it ever had.

\section*{The Witch's Closing}

``Context is a key. Lose the key, lose the door. Lose the door, you are still in the room. The room does not mind. The room is patient.''

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Hungry Scroll}

An ancient Lethai artifact that erases memory. Appears as a blank parchment in a sealed case. The Lethai do not destroy such things. They shelve them. The shelves have no labels.

\paragraph{Tags / Mechanics:}

\begin{itemize}
ℹtem Tags: [CURSE] [KNOWLEDGE] [DREAD]
ℹtem On reading, test Spirit + Resolve (DV 5). Failure: lose one memory (player chooses a minor fact −−− the name of a tavern, a route home, the sound of a friend's laugh).
ℹtem Timer : \textit{Erasure} [8]. When full, victim forgets their own name and cannot form new long−term memories. They become a Still−Room keeper. The job has no shortage of applicants.
ℹtem Destroyed only with fire from a living witness (a bell, a lamp, a human breath) −−− and the witness must speak the victim's true name while the paper burns. Lethai consider this rude. They do not say why.
\end{itemize}

\paragraph{Adventure Hook:}

A Lethai scholar has gone missing. Her notes are contradictions. In her study, a blank scroll on a lectern. The lamp is still warm. Read it, burn it, or seal it −−− knowing someone else will find it. Knowing the Still−Room has a space on the shelf.

\section*{Colophon}

Copied from a margin note in a book that was not supposed to leave Thepyrgos. The archivist would not let me see her face. The ink was still wet. She said, \emph{``Take it. I have forgotten why I kept it.''}

\noindent \textit{The paper is birch. The lamp flickers. The name Miralis is a placeholder.}