\chapter{The Bargain of the Shed Skin}

\section*{Told by}
Serpent−keeper Ashai, a dark elf with scales tattooed up her forearms and a voice like a knife wrapped in silk — soft until it is not. She has shed three names already. She remembers the first two.

\section*{The Tale}

A woman came to the Coil Court. She stood in the threshold, one boot on each side. "I gave a Lethaiar my fear of the dark. Now I cannot sleep. The faces in the shadows are not mine."

The serpent−queen did not look up from her coils. "You paid for a shedding. You received it. The price was never your fear. The price was your solitude."

The woman wept — a dry, rattling sound. "I did not know."

"You did not ask. The Coil Court does not sell answers. It sells skins."

The woman was quiet for a long breath. Then she offered a second bargain. "Shed this solitude. Forget the faces. Forget my mother's lullaby. I will owe a favor to the Court."

The serpent−queen raised her head. "A favor without limit?"

"Without limit."

The queen smiled. It was the smallest possible smile, the kind that says \textit{the hook is set}. She whispered a word. The woman shed.

She slept. Dreamless. Silent. For the first time in a year, she woke without a name on her tongue.

A year passed. Then a messenger came — a child with no shadow, carrying a folded leaf.

"The Court calls," the child said. "A child has been born without a shadow. You will nurse it until the shadow returns."

The woman nursed. One year. Three. Seven. The shadow never returned. The child grew — strange, calm, watchful, always polite. Never laughed. Never cried. When the child's mother died of a fever, the child shed its own name and joined the Coil Court.

The woman was free. The favor was paid.

But she often looked at her hands in the firelight and wondered whose heartbeat she was counting. Not hers. Not the child's. Something older.

She still cannot sleep. The solitude is gone. But the faces in the shadows have returned — and now she knows their names.

\section*{The Witch's Closing}
"A shedding does not erase the debt. It only changes the skin you wear. The scales remember. The shadows keep count."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Shadowless Child}

A child born without a shadow — an old bargain unpaid, a debt that has grown teeth. The child is unnervingly calm. Immune to fear. Can walk thresholds that bar others — doorways without permission, wards without invitation, the edge between sleep and waking. The child may be a vessel. The child may be the debt itself.

\paragraph{Tags / Mechanics:}
\begin{itemize}
ℹtem [THRESHOLD] [DREAD] [MARK]
ℹtem Immune to fear effects. Cannot be compelled, charmed, or intimidated by magic that targets emotion.
ℹtem Suffers −1 Position on Stealth rolls near the child. Shadows grow thin. Witnesses remember.
ℹtem Cannot be harmed by any creature that casts a shadow. Convert all incoming Harm to Fatigue (the shadow absorbs the wound).
ℹtem Timer : \textit{Return the Shadow} [8]. Ticks each session the child remains shadowless. When full, the original bargain finds a new debtor — possibly one of the characters.
\end{itemize}

\paragraph{Adventure Hook:}
A village is terrorized by a child who never sleeps, never blinks, and knows everyone's secrets — the ones whispered, the ones hidden, the ones not yet spoken. The child has no shadow. A Lethaiar came three years ago and claimed it as payment. Now the Lethaiar wants it back, and the child does not want to go.

\section*{Colophon}
Recorded in a Coil Court antechamber, with warm milk and bitter herbs. The lamp flickered once — no draft, no dying wick. The serpent−queen did not look at me once. When I left, she whispered a single word. I will not write it here.