\chapter{The Selkie's Oath}

\section*{Told by}
Captain Rana of the \textit{Eel−Skiff}, a Linn woman with salt−bleached braids and a laugh like shingle in a storm. Never goes ashore without a knife. Never sleeps without the hatch cracked.

\section*{The Tale}

A selkie married a fisherman. Same bargain as always: hide the skin, she stays. Lose it, she leaves.

He hid hers under the hearthstone. Ten years. Children. Mended nets. The quiet arithmetic of storms survived.

Then the youngest coughed blood into the porridge.

The nearest healer was a day's sail. The fisherman weighed the distance against the skin under the stone. His wife watched him do it.

She did not speak. She reached under the hearthstone. Pulled out the skin. Cut it into strips with his own gutting knife.

"Bandages," she said. "The child's breath is worth more than my freedom."

She wrapped the small chest. The fever broke before dawn. The skin was ruined — salt−stiff, chopped, fit only for the tide.

At first light, the fisherman touched her wrist. "Will you leave?"

She did not pull away. "I have no skin to return to."

"That is not an answer."

She turned his hand over, palm up. "I have a new oath. I choose to stay. That is heavier than any skin."

They say she grew webbing between her fingers the next winter. Her eyes turned the color of the deep — not grey, not blue, but the shifting place where light fails. The children learned to swim before they could walk.

The fisherman never asked again.

\section*{The Witch's Closing}
"Oaths change. So do the skins we wear. The sea does not care which — only that you choose, and that you choose while there is still breath."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Cut Skin}

A ruined selkie skin, cut into bandages, now stiff with salt and rust−colored stains. It heals any wound. It also remembers what it was.

\paragraph{Tags / Mechanics:}
\begin{itemize}
ℹtem [HEAL] [CURSE] [TRANSFORM]
ℹtem Heals 1 Harm per use. The user marks 1 Fatigue after each use — not from effort, but from hearing the sea call. The call gets louder each time.
ℹtem Timer : \textit{Sea−Change} [4]. Ticks each time the bandages are used. When full, the user gains a permanent minor mutation: webbed digits, salt−crusted skin that never dries, or eyes that reflect moonlight even indoors. No magic removes it short of a full renaming ritual at a crossroads at midnight.
ℹtem Once the timer fills, the bandages become inert. They have given everything.
\end{itemize}

\paragraph{Adventure Hook:}
A healer's bandages cure any sickness — but her patients wake dreaming of drowned cities, of stone towers under green water, of something that breathes in the trenches. The source is a cut selkie skin, kept in a locked chest. Destroy it and lose the miracle. Keep it and learn what the sea remembers.

\section*{Colophon}
Told on the deck of the \textit{Eel−Skiff}, rain coming in from the northeast. The captain spat over the rail. "That's the truth. The rest is embroidery."