\chapter{The Censor's Mask}

\section*{Told by}
Old Censor Varo, a man with one eye and a brand on his cheek — not punishment, he says, but \textit{documentation}. He sits by the forum fountain and feeds the pigeons. Children dare each other to ask him about the mask. He always answers. That is how you know he was the worst of them.

\section*{The Tale}

The Censor of Ecktoria wears a mask. Bronze, old, the eyeholes dark as a well at noon. It does not cover the face. It \textit{replaces} it. He wears it only at the Triumph, when the city brings its petitions and its lies.

One year, a merchant offered a bribe. A bag of silver. The Censor refused. The merchant doubled it. The Censor refused again.

"Name your price," the merchant whispered.

The Censor removed the mask.

The merchant screamed. There was no face underneath. Only a hollow space, and in that hollow, a whisper: \textit{I have seen your ledgers. Every page. Every missing grain.}

The merchant fled. The Censor put the mask back on. At the next Triumph, he wore it again.

But that night he dreamed of his own lies — the grain he had skimmed from the public store, the witness he had ignored because the bribe was smaller but sweeter. The mask had seen him too. The mask saw everyone. It simply waited.

He tried to remove it. The straps would not loosen. He cut them with a knife. The mask stayed. He heated the bronze. It cooled around his jaw. He broke a piece off with a hammer. By morning, the piece had grown back.

For a year he wore the mask. He could not eat. He could not drink. He did not starve. The mask fed on his guilt. He felt the hunger but could not die from it — suspended between confession and silence, like a fly in amber.

At the next Triumph, he walked to the forum. He knelt before the assembly. He said, "I am unworthy. I have stolen. I have lied. I have worn this mask as armor when it was meant to be a mirror. Judge me."

The mask fell off.

His face beneath was not old. It was young. His missing eye had grown back. The brand on his cheek was gone — erased, as if the skin had forgotten the iron.

He served thirty more years. The most honest Censor the city ever had. He never wore the mask again. He did not need to. The mask had taught him where to look.

\section*{The Witch's Closing}
"Judgment begins at home. The mask only shows what you already know — but will not say."

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Censor's Mask}

A bronze mask that shows the wearer their own hidden crimes — not visions, not revelations, just the quiet certainty of a conscience the wearer thought they had buried. It cannot be removed until the wearer confesses publicly. While worn, the wearer cannot eat or drink but does not die. Guilt is its own kind of bread.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [TRUTH] [PRESSURE]
\item Wearer gains +2 dice to Insight against any deception — the mask sees through lies because it knows the shape of a lie from the inside.
\item At the end of each scene where the mask is worn, the wearer must confess a secret flaw to another person. The GM may ask: \textit{What guilt does the mask show you right now?} If the player refuses, the clock \textit{Confession} [4] ticks.
\item When the clock fills, the wearer is compelled to make a full public confession in the next forum or marketplace they enter. The compulsion is absolute. After confession, the mask can be removed.
\item A character who has worn the mask for at least a year (in-game) and then confessed gains the Condition \textit{Unmasked}: +1 die to resist any future lie-based magic. They have seen the bottom of that well.
\end{itemize}

\paragraph{Adventure Hook:}
A city official has not eaten in three weeks. He speaks in whispers. A bronze mask has fused to his face. The party must learn what he did — and persuade him to confess before his body forgets how to want. The mask will not kill him. But it will wait.

\section*{Colophon}
Forum fountain, late winter, the light thin as old milk. The old man threw a crust to a pigeon. "I was his clerk," he said. "I forged the ledgers. He wore the mask for both of us. That was the bargain." He touched his cheek — the branded side. "I got the easier half."