\chapter{The Bride of Nine Knots}

\section*{Told by}
Lady Ysoria, a Vhasian widow who lives in a red tower and receives visitors only at dusk. Her left hand has six fingers. She says the extra one was a wedding gift. She did not say who gave it.

\section*{The Tale}

A bride tied her dowry rope with nine knots. One for each year she waited for her betrothed while he fought wars in the southern passes. The priest said, \textit{``Eight is custom. Nine is a door. Do not tie what you cannot untie.''}

She tied it anyway.

The morning of the wedding, her rival — who had also loved the groom, who had also waited, whose own dowry rope had only seven knots — stepped forward and cut the ninth knot with a blade of bone.

The groom forgot his vows. Not slowly. Not gently. He stood at the altar with his mouth half open, his eyes the color of a winter river, his hands empty. He did not reach for the bride. He did not speak. He simply \textit{stopped}.

The bride did not weep. She stepped around him. She took his hand — cold, still — and spoke for him.

\textit{``I give this ring. I take this name. I close this door. He is not here to refuse. That is not my problem. That is yours.''}

She said this to the priest. The priest was afraid. He nodded. The bond held.

But the rival's hand withered where she had cut the cord — five fingers turned to three, the skin drawn tight as dried leather, the bone showing at the knuckles. She wore a glove for the rest of her life. She never explained why.

The bride kept the severed knot in a silver locket. She wore it for fifty years. She never opened it. She never needed to.

On her deathbed, she gave the locket to her daughter.

\textit{``If you ever need to silence a liar,''} she said, \textit{``cut a knot. But be sure you know whose tongue will still speak afterwards. The blade does not ask whose blood it spills.''}

The daughter opened the locket. The knot was gone. Inside was a single tooth — the rival's, yellowed, cracked.

The daughter closed the locket. She did not ask.

\section*{The Witch's Closing}

\textit{``The ninth knot is the thread that holds the story closed. Cut it, and the story unravels — but not always in the direction you expect. The blade remembers who swung it.''}

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Severed Knot}

A piece of wedding cord, kept in a locket or a box or a hollow tooth. It remembers the vow it once bound. When cut again — for it can always be cut again — it does not destroy the vow. It simply transfers the forgetting. Someone else forgets. Someone else pays.

\paragraph{Tags / Mechanics:}
\begin{itemize}
ℹtem [CURSE] [TRUTH] [OATH]
ℹtem As an action, cut the knot and speak the name of a vow you witnessed. The person who made that vow forgets they ever spoke it. They will not remember until someone reminds them — and the reminder is always painful.
ℹtem The cutter chooses who forgets. The cutter does not choose who pays. The GM chooses a cost: a withered hand (−−1 die on physical rolls for one moon), a forgotten face (−−1 die on social rolls for one moon), or a debt transferred (an enemy becomes a friend; a friend becomes a stranger).
ℹtem The knot can be cut once per season. It grows back smaller each time. The ninth cut unravels the cutter's own name.
\end{itemize}

\paragraph{Adventure Hook:}
A noble has been forced into a marriage contract she does not want. She possesses a ninth knot — her grandmother's, passed down through three widows. Cutting the knot will make her betrothed forget his claim. But her grandmother's journal warns: \textit{``The third cut takes a finger. The sixth takes a memory. The ninth takes everything.''} The knot has been cut six times already. The party must decide: cut again, find another way — or ask whose debt they are really paying.

\section*{Colophon}
Red tower, dusk. The widow poured wine into eight cups. The ninth she emptied on the floor. \textit{``For the ones who forgot,''} she said. \textit{``And for the ones who will.''} The floor did not stain. The wine was older than the house.