\chapter{The Lantern That Remembered}

\section*{Told by}
Pilot Seref, a woman with one eye and a brass hook for a hand. She captains a little cutter called the \textit{Red Wake}. Says the sea tells her where to go. Says the sea is a poor liar but an excellent witness. When asked how she lost the eye, she taps the hook against her temple and says, \textit{"I looked at something that did not want to be seen."} She does not say what. The sea does not say either. The sea keeps its own accounts.

\section*{The Tale}

A young pilot lit a beacon on a headland that had no name — or rather, it had too many names, and the locals could not agree which one was true. His father had lit it before him. His grandfather had built it. The stone was warm even in winter, and the wind never blew quite the same direction twice.

It was his first night alone. His father had died the week before — not at sea, but in his bed, with the taste of salt on his lips and a question he had never asked. The young pilot did not know what the question was. He only knew that his father had stopped speaking halfway through a sentence, and the sentence had never finished.

He trimmed the lamp. He checked the mirror. He poured oil from a jug that had belonged to his great-grandmother, who had been a thief and a saint and possibly the same person. The flame burned steady. It was a small flame, no larger than a child's fist, but it threw light across the channel like a net.

At midnight, the beacon spoke.

Not in words, exactly. In a feeling — a pressure behind the eyes, a warmth at the base of the skull, a sense that something \textit{old} was turning its attention toward him in the same way a tide turns toward a gap in the rocks.

\textit{"I have seen this coast for three hundred years,"} it said. \textit{"I remember every ship that broke. I remember every pilot who listened. I remember the ones who did not."}

The young pilot's hands did not tremble. His father had taught him that. \textit{"The light does not fear the dark,"} his father used to say. \textit{"The dark is just the light's older sister. They argue. They always make up."}

"What do you want?" the pilot asked.

A pause. The flame flickered — not from wind, but from something inside the glass.

\textit{"I want someone to listen. Then I want someone to change the light. The channel shifts. The wrecks move. I cannot move myself. I am stone. I am patient. I am also tired."}

The pilot listened. He heard of wrecks he had never seen — ships that had sunk before his grandfather was born, crews that had drowned singing, a child who had floated ashore in a bread trough and grown up to be the pilot who replaced the pilot who had not listened. He heard of storms that came from nowhere and storms that came from everywhere. He heard the names of dead men spoken in a voice that was not quite a voice.

Then he climbed the tower. The ladder was cold. The rungs were worn smooth by the feet of a hundred years. He reached the lantern and turned it one degree east.

The beacon fell silent. Not an angry silence. A satisfied one. The silence of a stone that has finally settled after a long rain.

The next morning, a ship that would have wrecked passed safely. The pilot watched it from the headland, its sails white against the gray water. He did not wave. He did not cheer. He simply — \textit{listened} — to the silence, which was not empty, which had never been empty.

He was promoted. They gave him a higher tower and a better lamp and a jug of oil that had not belonged to his great-grandmother. He never told anyone about the voice. But every time he lit a new beacon — on a new headland, in a new channel, for a new generation of sailors who would never know his name — he spoke to it first.

\textit{"I am listening,"} he said. \textit{"Tell me when to move. Tell me what you remember. Tell me the names of the ones who did not listen, so I do not become one of them."}

And the beacons — some of them — answered.

\section*{The Witch's Closing}
\textit{"Beacons are not marks. They are arguments. A mark says, 'You are here.' An argument says, 'You should be somewhere else.' If you do not listen, the beacon will argue without you. And the sea will agree with the beacon. The sea always agrees with the oldest voice in the room."}

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Speaking Beacon}

An old lighthouse whose flame is alive. It remembers every ship that has wrecked nearby. It can see storms days before they arrive — not because it looks, but because it \textit{remembers} the last storm and the one before that and the one before that, and storms, like people, repeat themselves. It will share its knowledge only with those who listen — and who are willing to adjust the lantern's angle by a few degrees, changing the safe channel. The channel is never safe for long. The sea is patient. The sea is also a liar.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [KNOWLEDGE] [LOCATION] [TRUTH]
\item A character who spends a night in the beacon and succeeds on Spirit + Lore (DV 4) gains a vision of the next storm's path — not a map, but a feeling, a pressure behind the eyes, a certainty that the wind will shift at dawn. Gain +1 die to navigation for the next leg.
\item The beacon may also reveal the location of a sunken wreck (a clue, a treasure, or a body that needs to be found). The beacon does not judge the treasure. The beacon only remembers.
\item If the lantern is moved without the beacon's consent — if the pilot turns the light without listening first — the beacon goes dark for a month. The pilot gains a reputation as "deaf," which is worse than being blind, because the deaf cannot hear the warnings, and the warnings are always given.
\item A beacon that has been silenced for too long may begin to \textit{forget}. A forgetting beacon is dangerous. It may show safe water where there are rocks. It may show rocks where there is safe water. It may show nothing at all.
\end{itemize}

\paragraph{Adventure Hook:}
A coastal beacon has gone dark. Not dark like an unlit lamp — dark like a dead eye, a stopped clock, a story whose teller has walked away. Ships are wrecking on a reef that was never there before, or perhaps it was always there and the beacon had simply stopped remembering it. The party must discover why the lighthouse is silent — and whether the beacon is hiding something, grieving something, or waiting for someone. The pilot's father never finished his sentence. The party might be the ones who finish it.

\section*{Colophon}
On the \textit{Red Wake}, at anchor, with a rising tide and a sky the color of old tin. The pilot spat into the waves — not from contempt, but from custom. Her mother had spat before every voyage. Her mother's mother had spat before every birth. Spitting was a kind of prayer, she said. A prayer that did not ask for anything. A prayer that simply said, \textit{"I am here. The water is here. That is enough."}

\textit{"She's been quiet since the war,"} Pilot Seref said, jerking her hook toward the darkened headland. \textit{"I think she's waiting for someone. I think she's been waiting for a long time. I think she's tired of waiting."}

She spat again. The tide answered. The tide always answers, even when the beacon does not.