\chapter{The Loom of Witness}

\section*{Told by}
Weaver Thalia, a woman with red-stained fingers and a loom that never stops. She weaves shrouds for the guild --- for the ones who die with secrets still warm in their mouths. She does not ask what the secrets are. The loom tells her anyway.

\section*{The Tale}

A dyer bought a bolt of silk from a merchant who had found it in a wrecked ship, in a chest that was not locked but would not open until the fourth try. The silk was white --- not the white of fresh snow, but the white of bone after the marrow has been scraped out. Heavy. Cool to the touch. The merchant said, \textit{``This silk remembers every shame it has witnessed. It has been listening for a hundred years. Weave with care. Or do not weave at all.''}

The dyer did not believe him. He was a practical man. He believed in dye-vats, in weights and measures, in the price of madder root on the spring market. He dyed the silk crimson --- three baths, twelve hours, the color of a bruise the day after the blow --- and cut it into lengths for the market.

But the silk would not leave the loom. It rewound itself at night. The threads knotted into faces --- faces the dyer did not recognize at first, then faces he did. The faces whispered. They did not shout. They did not accuse. They simply \textit{told}. Confessions the dyer had never heard, shames he had never witnessed, weights he had never placed on his own scales.

The dyer tried to burn it. He built a fire in the yard, hot enough to melt bronze. The fire turned the silk black --- the black of a mouth that has been screaming for an hour and has nothing left --- and the whispers became louder. Not words now. Just sound. The sound of a hundred people trying to speak at once and finding no air.

A witch came. She did not knock. The door opened for her, or perhaps it had never been closed. She touched the loom --- not the silk, not the threads, but the wooden frame, worn smooth by fifty years of work.

\textit{``You did not buy the silk,''} she said. \textit{``You bought the witness. It has been following you since the merchant's stall. It will not leave until you confess.''}

The dyer knelt. His knees cracked. The floor was cold. He admitted his shames --- the false weights, the watered dyes, the apprentice he had blamed for his own mistake, and one more thing, smaller than the others, the thing he had almost forgotten: the cup of wine he had spilled on a customer's gown and not paid for, forty years ago, in a town whose name he could no longer remember.

The silk loosened. Not all at once. One thread at a time, like a fist opening.

The witch cut a single thread --- not from the silk, but from the air above the loom. It glowed for a moment, then dimmed.

\textit{``This is your toll,''} she said. \textit{``Wear it as a bracelet. When you lie, it tightens. When you confess, it loosens. When you die, it will return to the loom. The silk is patient. The silk has time.''}

The dyer wore the thread for the rest of his life. He never lied again --- not because he was virtuous, but because the thread made lying uncomfortable. A small discomfort. The discomfort of a too-tight shoe, or a collar that rubs. Just enough to remind him. Just enough.

The silk was sold for a fair price --- and the buyer was told its nature. The buyer was a temple. The temple hung the silk behind the altar. Confessions offered before it, the priests said, were heard more clearly. The silk listened. The silk remembered. The silk did not judge. That was the priests' job.

\section*{The Witch's Closing}

\textit{``What you hide may hold you. What you confess may free you. But freedom is not the same as forgetting. The silk remembers both --- the lie and the confession. It does not choose sides. It only weaves.''}

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Witness Silk}

A bolt of silk that retains memory of every wrong witnessed in its presence. It cannot be destroyed by normal fire --- fire is a confessor, not a judge, and the silk has already confessed everything it knows. To unbind it, someone must confess a secret shame in front of it. The shame does not need to be large. It only needs to be true. The confession frees one yard of silk at a time. The freed silk is inert --- a cloth like any other. But the yard that remains keeps listening.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [TRUTH] [WITNESS] [BURDEN]
\item A character who possesses the silk gains +1 die to Insight rolls (the silk whispers hints --- not clear answers, but textures, feelings, the shape of a lie). However, at each downtime, the GM may ask for a confession --- a minor flaw, a forgotten slight, a small cruelty. If the character refuses, the silk tightens: --1 die on all social rolls until they confess. The tightening is physical. The character can feel it.
\item The silk can be ``paid off'' by sacrificing a yard (which becomes inert cloth) in exchange for a favor from a guild or spirit. The favor is never large. The silk is not generous. But a small favor --- a door left unlocked, a message delayed, a name forgotten --- is sometimes enough.
\item The silk has a number of yards equal to the number of confessions it has witnessed. Every new confession adds a yard. Every sacrifice removes one. The silk can never be reduced to zero. It always keeps at least one yard --- the yard that witnessed its own creation, in a wrecked ship, in a chest that would not open.
\end{itemize}

\paragraph{Adventure Hook:}
A merchant's best loom is stuck --- the silk will not come out. The loom has been in her family for six generations. Her grandmother wove the shrouds for the plague dead. Her great-grandfather wove the sails for a ship that never returned. The silk knows something about each of them --- something none of them confessed. The merchant is hiding a secret that could ruin her. Not a large secret. A small one. The kind that fits in a cup of wine, spilled forty years ago, in a town whose name she can no longer remember. The party must decide: extract the confession, burn the loom, or find a third way --- knowing that the silk will remember whichever path they choose.

\section*{Colophon}
Weaver's loft, dusk. The loom clicked --- not a sound of work, but a sound of listening. Thalia touched a thread. Not the crimson silk. A different thread, gray, thin, almost invisible against her skin.

\textit{``This one is mine,''} she said. \textit{``I overcharged a widow. Twenty years ago. The widow is dead now. The overcharge was three coppers. The silk does not care about the amount. It only cares that I remember.''}

She did not cut the thread. She let it be.

The loom clicked again. Thalia did not look up.