\chapter{The Coin That Came Back}

\section*{Told by}
Old Salt, a Zakov pirate with a peg leg and a parrot that swears in three languages. He has been drowned twice. The sea spit him back both times, but the second time it kept his left ear. He says the ear still hears underwater. He says it hears things no living person should know. He will not say what.

\section*{The Tale}

A smuggler found a coin in a dead man's eye. Not pressed into the socket---resting there, warm as a fresh egg, as if the dead man had been saving it for someone. The coin was silver, stamped with a ship that had no name, its sails furled, its masts bare. The smuggler pried it loose with two fingers. The eye did not close. The dead man did not stop looking.

He took it. A smuggler who will not take a coin from a corpse is a smuggler who has never been hungry.

That night, the coin bought him safe passage past a patrol. The watchman looked at it---just looked, just a glance---and waved him through without a word. Without a toll. Without a name. The watchman did not remember the encounter. Later, when asked, he said he had no memory of any ship passing that night. He said he must have been asleep. He said he never slept on watch.

The next day, the dead man's ghost sat in his cargo hold. Not floating. Not wailing. Sitting, cross-legged, on a barrel of salted herring, its hands folded in its lap, its eye sockets empty but somehow still watching.

\textit{``You took my toll,''} the ghost said. Its voice was the sound of water finding a crack in a hull. \textit{``Now you owe me passage.''}

The smuggler tried to sell the coin. No one would take it. Merchants looked at it and turned pale. Pawnbrokers refused to touch it. A priest of the sea-god held it in his palm, listened for a long moment, and handed it back without a word. His hand was shaking.

He tried to throw it overboard. It floated back---not drifting with the current, not washing up on the tide, but rising straight up through the water like a bubble, landing on the deck at his feet with a sound like a bell.

He tried to give it to a beggar. The beggar returned it at dawn, pale as milk, and would not say where he had found it. He only said, \textit{``It followed me. It knows your name.''}

He sailed to a shrine of the drowned---a stone chapel built on a skerry, reachable only at low tide, its floor covered in salt and broken wishes. He placed the coin on the altar. The ghost appeared again, sitting on the altar now, its legs dangling over the edge like a child waiting for a parent.

\textit{``Not enough,''} it said. \textit{``The shrine is not where I died. You must carry me to the water that took me. You must carry me to the reef. You must carry me home.''}

The smuggler sailed to a reef---a place of broken ships and bleached coral, where the water was the color of bruises. He dropped the coin into the water. It did not float back. It sank, slow as a dying star, and the ghost smiled---a thin smile, a relieved smile, a smile that said \textit{finally}---took the coin, and vanished.

The smuggler sailed home. His hold was full of fish---more than he had ever caught, more than his nets could hold, fish of species he had never seen, fish with scales the color of sunset and eyes the color of the coin. He sold them. He grew rich. He never picked up a coin from a dead man's eye again.

But sometimes, on still nights, he hears a coin spinning on the deck of his ship. He does not look. He knows better. The dead have long memories, and the sea has a longer one, and a coin that has been returned is not a coin that has been forgotten. It is a coin that is waiting.

\section*{The Witch's Closing}
\textit{``Not all treasure is worth its weight. Some coins carry a ticket. The ticket is always for a destination you would not choose. The sea does not ask whether you want to go. The sea only asks whether you are ready to pay the fare.''}

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Ferry Coin}

A silver coin taken from a corpse's eye. It cannot be spent, given away, or destroyed---not by fire, not by water, not by prayer. It follows. It waits. It knows your name. It forces the holder to carry the ghost of the dead person to the place of their death. Not the place they were buried. Not the place they were born. The place where the water took them. Once there, the coin and ghost vanish, and the holder is granted a boon of the sea: good fishing, calm waters, a safe passage, or---if the ghost is grateful---a single answer to a question about drowning.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [CURSE] [SACRIFICE] [TRUTH]
\item The holder gains +1 die on all Negotiation rolls (the coin whispers advice---the dead know things the living have forgotten) but suffers nightmares every night unless they are actively sailing toward the ghost's destination. The nightmares are not vague. The holder dreams the ghost's death, over and over, from the inside. Each night, the holder wakes with salt on their lips and water in their lungs.
\item The ghost can be negotiated with---it might accept a substitute voyage (delivering a message, carrying an object, telling a story to a specific person) instead of a literal journey. This requires a Spirit + Lore test (DV 5). On a failure, the ghost does not refuse. It simply stares. The stare is worse than any refusal.
\item The coin cannot be destroyed. It can be hidden---sealed in lead, buried in dry earth, dropped into a volcano. It will return. It always returns. The only way to be free of it is to pay the passage.
\end{itemize}

\paragraph{Adventure Hook:}
A sailor has a coin that will not let him sleep. He does not know whose ghost it is---only that the ghost is old, and patient, and has been waiting a very long time. The party must help him discover the dead person's identity and their destination before the ghost stops waiting. Because when a ghost stops waiting, it starts taking. First the sleep. Then the memories. Then the name. And then---the coin finds a new pocket.

\section*{Colophon}
Zakov tavern, the \textit{Drowned Rat}. Old Salt flipped a coin to the barkeep. \textit{``That is a live one,''} he said. \textit{``Keep the change. Or don't. It won't matter. The coin knows where you live.''} The barkeep put the coin in the till. The next morning, it was on the bar again. The barkeep has not slept since.