\chapter{The Ford That Moved}

\section*{Told by}
Ferryman Dav, a Dhaharan with a turban and a voice like gravel rolled downhill. He poled his boat across the same ford for fifty years. He says the river changed course three times in his lifetime—once after a flood, once after a famine, and once after a wedding where the groom forgot the bride's name halfway through the vows. He does not know which change was which. He only knows the river has moods, and the moods have teeth.

\section*{The Tale}

A caravan master needed to cross a ford. The river was high, the stones slick as oil, the current muttering to itself in a language that sounded like breaking glass. He had crossed this ford a hundred times. This time, it felt different. This time, the water was watching.

He knelt and prayed to the ford—not to the river, not to the god of rivers, but to the ford itself, the place where water met stone and stone let travelers pass. He offered a single grain of wheat, held between thumb and forefinger, small as a promise.

\textit{``I will stay still for one crossing,''} the ford whispered. The voice came from nowhere and everywhere—from the water, from the stones, from the air just above the surface. \textit{``Then you will owe me a name.''}

The caravan master crossed. The ford held. The water did not rise. The stones did not shift. He reached the far shore with dry boots and a wet back—his own sweat, not the river's.

The next night, he returned. The river was different—deeper, swifter, darker. The ford had moved. The crossing was a hundred paces downstream, and the old path was now a whirlpool the size a horse could drown in.

\textit{``Name a price,''} the merchant said.

\textit{``You have already named it,''} the ford replied. Its voice was softer now, almost kind. Almost. \textit{``The name of your firstborn child. The name you whisper when you think no one is listening. The name you have kept folded in your chest like a letter you are afraid to send.''}

The merchant had no child. He was childless. His wife had died in the winter. He had buried her with his own hands, and he had not wept then, but he wept now—great heaving sobs that echoed off the water and came back to him twisted, mocking.

The ford softened. The current slowed. The stones rose just enough to offer a path.

\textit{``Then give me the name you would have given them. The name that died in your throat when your wife died. The name that has no grave, no marker, no witness. Give me that name. And every year, on this night, return and speak it again. That is the toll. That is always the toll.''}

The merchant knelt. He pressed his forehead to the wet stones. He whispered a name—a name no one had ever heard, a name he had never spoken aloud, a name that existed only in the space between his heartbeat and his breath.

The ford calmed. The water stilled. He crossed safely, and the far shore rose to meet him like an old friend.

He returned every year. He spoke the name. The ford accepted it, drank it, swallowed it whole. His wife did not come back. The child did not appear. But the name—the name stayed alive, spoken once a year into the water, and the water carried it somewhere the merchant could not follow.

His wife gave birth to a daughter. He named her the same name. The ford never troubled him again. But the daughter—the daughter sometimes woke from dreams she could not remember, with salt on her lips and the sound of water in her ears. She never learned to swim. She never wanted to.

\section*{The Witch's Closing}
\textit{``A road is a covenant. A ford is a witness. The river remembers every name it has been given. It does not forget. It does not forgive. It only waits for the next traveler who thinks a grain of wheat is enough.''}

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Naming Ford}

A river crossing that asks for a name as toll. Not a coin. Not a prayer. A name. It will accept a false name once—once, and only once, and only if the false name is beautiful enough to make the water laugh. The next crossing demands the true name of the traveler. Not the name on their papers. Not the name their mother gave them. The name they whisper to themselves in the dark, when they are afraid, when they are alone, when they are not sure they deserve to exist.

If the traveler refuses, the ford shifts its course—sometimes by a few paces, sometimes by a mile, sometimes by a generation. It becomes impassable for a year and a day. By then, the traveler is usually dead, or lost, or someone else entirely.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [THRESHOLD] [TRUTH] [SACRIFICE]
\item To cross, a character must succeed on a Spirit + Lore test (DV 3) to recall a name that has meaning—not necessarily their own, but a name that cost them something to learn. On failure, the ford demands a true name from one of the party. That character gains a String on themselves (the ford knows their name, and the river remembers). Water magic targeting that character gains +1 die for the next moon. The ford does not explain this. The ford does not need to.
\item If the party tries to cross without paying, the ford becomes a hazard: Athletics DV 5 to navigate the shifting stones and sudden currents. On failure, the character is swept downstream, loses 1 Supply, and arrives at the far shore with wet boots and a new understanding: the ford was being polite before. It has stopped being polite.
\item A character who pays the ford with a true name may, once per session, ask the GM: \textit{``What does the river know that I do not?''} The GM answers with a single fact about the nearest body of water—a hidden current, a drowned body, a secret the water has kept. The answer is always true. The answer is never comforting.
\end{itemize}

\paragraph{Adventure Hook:}
A ford has closed to all trade. The local lord sent a surveyor; the surveyor returned, but he cannot remember his own name. He knows his face. He knows his horse. He knows the way home. But when someone asks \textit{``who are you?''} he opens his mouth and nothing comes out—only the sound of water, distant and cold. The party must learn what the ford took—and whether a name, once given, can ever be returned. The river says no. The river has been wrong before. Not often. But once is enough.

\section*{Colophon}
Ferryman's hut, near the saffron fields. Dav spat into the water. The water spat back. \textit{``She is quiet tonight,''} he said, wiping his cheek. \textit{``Must have taken a good one. A ripe one. A name with weight.''} He poled his boat into the current. The river did not move. It was waiting. It is always waiting.