\chapter{The Counting of the Fifth Breath}

\section*{Told by}
Sergeant Relled, a Black Banner mercenary with a missing ear and a habit of counting his steps. He has survived seven campaigns. He does not know why. He has stopped asking. The asking, he says, is what gets you killed. The counting keeps you alive.

\section*{The Tale}

A young captain lost his nerve in a skirmish in the Singing Wastes. Not his sword arm. Not his voice. His nerve. The kind that sits behind the ribs and tells the hands what to do when the mind has frozen. The enemy had five breaths to plant a banner on the ridge. Five heartbeats. Five chances. Five ways to lose everything.

The captain panicked. Not the dramatic panic of a screaming recruit. The quiet panic. The one where the eyes go wide and the mouth opens and nothing comes out.

\textit{``I have four breaths left,''} he told his sergeant. His voice was the sound of a blade scraping a stone. \textit{``What do I do?''}

The sergeant did not panic. The sergeant had survived seven campaigns by not panicking. He looked at the ridge. He counted. Not the breaths. The exits.

\textit{``How many ways up?''}

\textit{``Two.''}

The sergeant nodded. \textit{``Then they are already dead. They just do not know it yet. Weld one path. Block the other. Use the first breath to drink water. Your mouth is dry. Dry mouths make bad decisions.''}

The captain drank. His hand stopped shaking.

The sergeant welded the path with a line of reds---not a wall, not a barricade, just a suggestion of fire, a promise of pain. The enemy runner hesitated. Hesitation is a kind of death. It is not quick. It is not clean. But it is final.

The captain's second breath, he moved an orange to fork the ridge. Not to attack. To divide. An enemy that must choose is an enemy that has already lost the advantage of speed.

Third breath, he planted a green on the low path. Not a soldier. A flag. A suggestion that someone was already there. The enemy runner stopped. He looked at the flag. He looked at the ridge. He looked at the burning path.

Fourth breath, the captain waited.

Waiting is the hardest breath. The lungs want to empty. The hands want to move. The heart wants to do \textit{something}---anything---even the wrong thing. The captain waited.

The enemy runner never reached the ridge. The fifth breath expired. The banner stayed in its scabbard. The captain won.

Later, the sergeant sat with him by the fire. The captain's hands were still shaking, but his voice was steady.

\textit{``How did you know?''}

\textit{``I did not know,''} the sergeant said. \textit{``I counted. Counting is not knowing. Counting is buying time to think. Panic is a thief. It steals the breaths you cannot afford to lose. Do not let it. Count the breaths. Then count your exits. Then count your water. Then move. In that order. Never skip the water. Dry mouths make bad decisions.''}

\section*{The Witch's Closing}
\textit{``Calm is not a gift. It is a habit. Spend your first breath on it. The rest will follow---or they will not. But at least you will have spent the first breath well.''}

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Five-Breath Timer }

A timer that appears when a character is in a desperate race---a siege, a chase, a ritual, a collapsing ceiling, a burning ship. Not a magical construct. Not a curse. Simply the shape of pressure made visible: five breaths, counting down. The character can see the count---not as numbers, but as a sensation behind the eyes, a tightness in the chest, a voice that whispers \textit{five, four, three} with every heartbeat.

The timer is not an enemy. The timer is a mirror. It shows what you are made of.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [PRESSURE] [DREAD] [TRUTH]
\item When the timer appears, the character gains +1 die to all actions directly related to the objective---panic focus, the body's cruel gift of narrow attention. But they suffer --1 die to all other actions. Tunnel vision. The enemy in the corner of the eye. The step they did not see.
\item If the timer reaches zero, the objective fails---but failure is not the end. Failure is a door. The GM may offer a Devil's Bargain: succeed anyway, but lose something else. A memory. A name. A breath you will never get back.
\item The character can reset the timer by finding a new approach---a side path, a different tactic, a sacrifice the enemy did not expect. This requires a Wits + Tactics test (DV 4). On a failure, the timer does not reset. It simply... watches. Waiting is the hardest breath.
\end{itemize}

\paragraph{Adventure Hook:}
A mercenary company is trapped in a collapsing fortress. They have five minutes before the enemy breaches the inner gate. The captain carries a magical hourglass---not one that measures time, but one that \textit{remembers} it. The hourglass ticks audibly, and with each tick, a grain of sand falls not into the bottom chamber, but into a third chamber no one has ever seen. The captain does not know what happens when that chamber fills. Neither does the sergeant. Neither does the enemy. The party must complete their escape---or find a way to break the hourglass before everyone discovers the answer together.

\section*{Colophon}
Campfire near the Singing Wastes. The wind carried voices from the dunes---old voices, hungry voices, voices that had been counting for centuries. The sergeant ignored them. He counted his fingers instead. \textit{``Five,''} he said. \textit{``Always check after a fight. You never know when one went missing. Fingers, ears, breaths. Check them all. Then drink water. Dry mouths make bad decisions.''} He drank. The wind stopped. The voices went elsewhere, looking for someone who would listen.