\chapter{The Bargain of the Unlit Lamp}

\section*{Told by}
A Trow who gave no name. It sat in the corner of the tavern, smoking a pipe that had no ember and no stem---just a bowl, cradled in gray fingers, empty but somehow still drawing. Its voice was dry leaves skittering across stone. Its hat drank the starlight. I wrote what I heard. When I checked my notes the next morning, the ink had faded. I rewrote from memory. The memory was not mine. I do not know whose it was.

\section*{The Tale}

A traveler came to a crossroads at midnight. Not the kind of crossroads where two roads meet---the older kind, where the path between the living and the hollow crosses itself, where the shadows lean the wrong way, where the wind smells of extinguished hearths. A figure waited there. It wore a gray robe and a hat that drank the starlight. Its face, if it had one, was hidden. Its hands, if it had them, were folded.

\textit{``I am a Trow,''} it said. \textit{``We trade in absence. We deal in the spaces where things used to be. What do you lack?''}

The traveler lacked nothing. He was wealthy, healthy, happy, loved. He had a wife who smiled when he entered a room. He had children who ran to meet him at the gate. He had a dog that slept on his feet. He had a house with a door that swung true and a roof that did not leak.

\textit{``Then sell me something you do not need,''} said the Trow. \textit{``A memory you have not used in years. A breath you were going to waste. A name you have never spoken aloud. A glove that has a match. A single heartbeat from the pile of thousands.''}

The traveler laughed. The laugh was small and thin. The crossroads did not echo. Nothing echoes at a Trow's crossroads. He pulled off his left-handed glove---a good glove, leather, warm, but he had another just like it---and held it out.

The Trow took the glove. It did not put it on. It folded it once, twice, three times, until the glove was the size of a coin, and then it tucked the folded glove into its sleeve. The sleeve did not bulge. The sleeve had been waiting.

The Trow gave him a lamp---a small thing, brass, unlit, with a wick that looked like it had never seen flame. \textit{``The light will not fade,''} it said. \textit{``It will burn forever. It will never need oil. But you will forget why you needed it. You will forget what you wanted to see. The light remembers. You will not.''}

The traveler took the lamp. He walked home. The light guided him---a steady, warm glow that showed every root, every stone, every shadow that might hide a hole. But when he arrived at his gate, he paused. The gate looked wrong. The paint was the right color, but the shape was unfamiliar. He opened it. The path to the door was the same length, but the door itself---the door was not his door.

He opened it anyway. Inside, a woman looked up from the hearth. Her face was kind. Her hair was gray. He did not know her.

\textit{``Who are you?''} he asked.

She stared. \textit{``I am your wife.''}

\textit{``I do not have a wife,''} he said. He meant it. He could not remember ever having a wife. He could not remember ever being loved. The lamp burned steadily in his hand, warm as a heartbeat, patient as stone.

He turned and walked back to the crossroads. The Trow was gone. The only thing on the ground was his glove---unfolded, full-sized, lying in the dust as if it had been dropped by someone in a hurry. He put it on. The lamp went out. The darkness was absolute. He sat in it for a long time, remembering.

He remembered his wife's face. He remembered her name. He remembered the way she laughed when he tripped over the dog. He remembered the dog. He remembered everything the lamp had taken, and he remembered that he had given it willingly, and he remembered that he had laughed.

He left the lamp at the crossroads. He walked home in the dark. He did not look back. The lamp was still there the next morning, and the next, and the next. No one took it. The Trow did not return.

He never bought from a Trow again. But sometimes, on winter nights, when the wind was right, he would stand at his gate and look toward the crossroads. He could not see the lamp. It was too far. But he could feel it---waiting, burning, remembering everything he had been willing to forget.

\section*{The Witch's Closing}
\textit{``Trow trade in gaps. The gap that light fills is the gap where memory used to be. Do not buy a lamp from a creature that wears a hat made of starlight. The hat is not the problem. The hat is just a hat. The problem is the seller. The seller does not need your glove. The seller needs your forgetting.''}

\section*{What Lurks in the Shadow of the Tale}

\subsection*{The Trow Lamp}

A lantern that never needs oil and never goes out---not for lack of fuel, not for lack of air, not for lack of time. It burns on absence. It feeds on what you no longer remember. Each hour it burns, the user loses a random memory---a face, a word, a route, a name, a scar, a song, the smell of rain on dry ground, the sound of a specific voice laughing. The light is perfect. The cost is everything. Once lit, the lamp cannot be extinguished except by returning the price paid---and the Trow will accept that return only once, only at the same crossroads, only at midnight. The Trow is patient. The Trow is always there.

\paragraph{Tags / Mechanics:}
\begin{itemize}
\item [LIGHT] [CURSE] [SILENCE] [SECRET]
\item The lamp provides perfect illumination in a 30-foot radius. No shadows. No blind spots. No hiding. The light does not flicker. The light does not judge. The light simply \textit{shows}.
\item Every hour of continuous use, the GM erases one minor fact from the character's background---the name of a childhood friend, the layout of a familiar city, the face of a parent, the way home. The character may still know the fact exists. They cannot recall it under pressure. They cannot find it when they need it. It is a hole in their head, warm and smooth and empty.
\item After 8 hours of cumulative use, the character cannot remember their own name. After 16 hours, they cannot remember why they are holding a lamp. After 24 hours, they cannot remember how to speak. The lamp burns on. The light is still perfect.
\item To extinguish the lamp, the character must give up something of equal value to what they originally paid---not a glove, this time, but something that costs. A cherished possession that cannot be replaced. A close relationship, burned and broken. A skill point, lost as if it had never been learned. A memory they were trying to keep. The Trow will know if the payment is false. The Trow always knows.
\end{itemize}

\paragraph{Adventure Hook:}
A lighthouse keeper has used a Trow lamp for thirty years. The light has never failed. Ships have never wrecked on his shore. But the keeper has forgotten his own daughter's name---and her face, and the way to the village, and the reason he ever wanted a lamp in the first place. He sits in the tower, staring at the flame, while the sea rises and the rocks grow sharper. The party must convince him to extinguish the light before the next storm---or find a way to pay the Trow's price themselves. The lamp is patient. The lamp is always patient. The storm is not.

\section*{Colophon}
Tavern corner, after midnight. The Trow smiled. Its teeth were gray as sea-worn stone, set in gums the color of old parchment. \textit{``I will take that question,''} it said. \textit{``You have already forgotten you asked it. The forgetting is the payment. The question is the interest. The interest is always due.''} I looked at my notes. The page was blank. The ink was gone. The lamp on the table---the one I had lit to write by---had gone out. I do not remember when. I do not remember why. I do not remember lighting it at all.