\chapter*{Colophon: Margins and Dust}

The paper is pressed flax---soft, cheap, good for ledgers, good for wrapping fish, good for the bottoms of birdcages. It was never meant to hold stories. It was meant to hold sums. Columns of grain. Counts of coin. Names of debtors, scratched out and rewritten, scratched out again. The stories are trespassing. The stories do not care.

Ink is lamp-black and oak gall, mixed with a drop of my own blood. Not for binding. For memory. Blood remembers what ink forgets. Blood has been here before. Blood will be here after. The oak gall came from a tree that witnessed a hanging. The lamp-black came from a fire that warmed a murderer. I did not know this when I mixed them. I know it now. The ink knows more than I do. The ink is older.

\smallskip
Ikasha asks for a secret with every telling. She does not demand. She does not threaten. She simply \textit{listens}, and the listening is the price. A secret told is a secret spent. You cannot spend the same secret twice. I have spent nine in this book. The ninth is not written. The ninth is the space between these words, the place where the eye slides over, the place where the page is blank and the blank is full.

I gave her the corners of these pages, where the words thin and the eye looks away. The corners are where she sits, curled like a cat, patient as stone. She does not turn the pages. She does not need to. She hears the ink drying. She hears the fibers settling. She hears the silence between the tales, and the silence is the part she came for.

\smallskip
If you find this book in a chest, on a ferry, or lying open in a shrine---leave a copper coin on the cover. Not for me. The toll is for the road. The road has been patient. The road has been waiting. The road has not forgotten the weight of your boots. A copper coin is light. The road does not mind. The road only asks to be acknowledged.

Do not leave silver. Silver is for the dead. This book is not for the dead. This book is for the living who are still learning how to owe.

\smallskip
The fires are cold. I have another leg to walk. The leg is not long---a day, maybe two, to the coast, to a ship that may or may not be there, to a captain who may or may not remember my face. The Tulkani do not pray for safe passage. We pray for interesting weather. Interesting weather makes good stories. Good stories pay the toll. The toll is always due.

\smallskip
\noindent --- \textit{Dusana}  

\noindent \textit{Crossroads east of Silkstrand, dawn of the autumn equinox.}  

\noindent \textit{I poured water first, then drank. The bell's shadow fell across my page. I left before it moved. The shadow is still there, I think, waiting for the next hand to hold the page. The shadow does not move. The shadow is patient. The shadow has been here before.}