\section{Dawn Warrior (Oath of Flame and Light)}

\subsection{The Story of Ser Roran}

\epigraph{He knelt at the bank of the Black River. The water was cold. It smelled of iron and old rain. He opened the locket.}{From \textit{The Black River at Dawn}}

Ser Roran was not always a Dawn Warrior. He was not always called Roran. Once, he was Aldric, son of Baron Aldric of the Scar—a boy with a father who laughed when asked to pay his ransom, a mother who died giving him life, and a wetnurse who lost her hand for a locket.

He spent three years in a debtor's cell. A stranger paid his ransom—a hedge-knight who had once served his father, who spent every coin he had and died of fever the following winter. The stranger gave Roran his name and his sword.

Roran knighted himself on a battlefield, with a broken blade and the blessing of a priest who witnessed his vow. He swore never to take a ransom. He swore to protect the helpless. He swore to remember his mother's face, even though he had never seen it.

His Codex is a silver gauntlet, engraved with the names of every person who has ever shown him mercy. His Thiasos is a warhorse with eyes like dying embers. His Rites are the vows he speaks before battle—each one a promise carved into steel.

\subsection{The Scar That Remembers}

Roran's left hand is sheathed in a silver gauntlet. Beneath it, the hand is whole. The gauntlet is not armor—it is a reminder. The stranger who ransomed him wore a gauntlet on his left hand. He died with it on. Roran never took it off.

The gauntlet is warm when he is keeping his vow. It is cold when he is tempted to break it. He has never felt it cold.

\subsection{The Vow That Unbinds}

At the trial of Baron Aldric, Roran did not speak against his father. He did not need to. He simply stood at the bar, the silver gauntlet on his left hand, and waited. When the Right of Acknowledgment was invoked, he did not demand a confession. He demanded only this:

\vspace{0.5em}
\begin{quote}
\textit{"Say: I was afraid."}
\end{quote}
\vspace{0.5em}

His father said it. The vow was broken. The knot was cut.

Roran walked out of the courtroom. He did not look back.

\subsection{The River at Dawn}

Years later, Roran carried his mother's locket to the Black River. He did not know that Bryn had kept it for forty years. He did not know that his mother's ghost was watching. He only knew that the locket was warm, and the thread was red, and the water was cold.

He opened his hand. The water rose. The locket sank.

When he looked up, the mist had cleared. The sun was rising. He did not look back.

\subsection{What the Dawn Warrior Carries}

A Dawn Warrior carries their vows in steel and fire. They carry the memory of every mercy they have received. They carry the weight of every oath they have sworn.

But most of all, they carry the knowledge that the strongest vow is the one that unbinds.

\vspace{1em}
\begin{center}
\textit{The thread held. The water carried. That is not a Rite. That is grace.}
\end{center}

\section{Dawn Warrior (Oath of Flame and Light)}

\subsection{Concept}

The Dawn Warrior is a crusader who carries their vows in steel and fire. They are not scholars or courtiers. They are the ones who stand at the front, who hold the line, who charge when others hesitate. Their Codex is an engraved gorget or gauntlet. Their Thiasos is a warhorse with eyes like embers or a spirit-hound of ash and seal-wax. Their Rites are not whispered—they are declared.

A Dawn Warrior's covenant is visible. Their vows are carved into their armor. Their scars are receipts for oaths kept. They do not hide what they are.

\subsection{Core Attributes}

\begin{itemize}
\item \textbf{Primary:} Body, Spirit
\item \textbf{Secondary:} Presence
\item \textbf{Tertiary:} Wits
\end{itemize}

\subsection{Skills}

\begin{itemize}
\item \textbf{Core:} Melee, Command, Athletics
\item \textbf{Secondary:} Lore (Oaths), Survival
\item \textbf{Tertiary:} Insight, Stealth
\end{itemize}

\subsection{Codex}

The Dawn Warrior's Codex is worn into battle:

\begin{itemize}
\item \textbf{Engraved Gorget:} A steel collar worn over the throat. Each vow is a line of flame-light script. The gorget grants +1 Armor.
\item \textbf{Branded Gauntlet:} A plate glove marked with runes of dawn. The hand that strikes also swears.
\item \textbf{Dawnblade (Sentient):} A sword that hums with trapped light. Speaks in proverbs about honor. Refuses to strike an unarmed foe. (See Optional Rule: Sentient Thiasos)
\end{itemize}

\subsection{Thiasos}

\begin{itemize}
\item \textbf{Warhorse:} A horse with eyes like dying embers. Fears no flame, will not bolt from battle. Cap 3, Harm 5.
\item \textbf{Spirit-Hound of Ash:} A hound whose fur smolders and whose paws leave no tracks. Can sense oath-breakers within Near range. Cap 2, Harm 4.
\item \textbf{Dawnblade (Unified Codex):} The sentient blade serves as both Codex and Thiasos (see Unified Codex optional rule).
\end{itemize}

\subsection{Signature Rites}

\subsubsection{Low Rite: Rite of the Kindled Blade (4 XP)}

Your weapon ignites with radiant light. For one scene, your melee attacks deal +1 Harm against undead, oath-breakers, and creatures of corruption.

\subsubsection{Standard Rite: Rite of the Unbroken Vow (7 XP)}

Swear an oath before a flame and a witness. For the duration of the oath, you gain +1 Armor. Any ally who hears the oath and aids you gains +1 die to actions that serve it.

\subsubsection{High Rite: Rite of the Dawnfire Covenant (12 XP)}

Consecrate a zone as dawnfire ground. Allies within gain +1 die to resist fear. Enemies who are undead, oath-breakers, or creatures of corruption suffer -1 die to all actions.

\subsection{Signature Talents}

\subsubsection{Dawn's Judgment (6 XP)}

When you invoke Rite of the Kindled Blade, you may choose to extend the effect. Your blade remains kindled for the remainder of the scene. Each additional attack beyond the first costs 1 Fatigue.

\subsubsection{Unquenchable (4 XP)}

Your flame cannot be extinguished by mundane means. Wind, water, and darkness do not affect your kindled blade or dawnfire Rites.

\subsection{Sample Build — Tier I (30-40 XP)}

\begin{itemize}
\item \textbf{Attributes:} Body 3 (9 XP), Spirit 2 (6 XP), Presence 2 (6 XP), Wits 1 (3 XP) — Total 24 XP
\item \textbf{Skills:} Melee 2 (4 XP), Command 1 (2 XP), Athletics 1 (2 XP) — Total 8 XP
\item \textbf{Talents:} Flesh Codex (2 XP), Oath-Bound (4 XP), Scarred Resilience (3 XP) — Total 9 XP
\item \textbf{Remaining XP:} 0 (30 XP base + bonds/complications as needed)
\end{itemize}

\subsection{Sample Build — Tier III (91-150 XP)}

\begin{itemize}
\item \textbf{Attributes:} Body 4 (18 XP), Spirit 3 (12 XP), Presence 3 (12 XP), Wits 2 (6 XP) — Total 48 XP
\item \textbf{Skills:} Melee 3 (9 XP), Command 2 (6 XP), Athletics 2 (6 XP), Lore (Oaths) 2 (6 XP), Survival 1 (2 XP) — Total 29 XP
\item \textbf{Talents:} Flesh Codex, Oath-Bound, Scarred Resilience, Engraved Defense (6 XP), Iron Resolve (7 XP), Dawn's Judgment (6 XP) — Total 19 XP (cumulative)
\item \textbf{Rites:} Rite of the Kindled Blade (4 XP), Rite of the Unbroken Vow (7 XP) — Total 11 XP (cumulative)
\item \textbf{Total XP:} 48 + 29 + 19 + 11 = 107 XP
\end{itemize}

\subsection{Roleplaying a Dawn Warrior}

\begin{itemize}
\item You do not whisper your vows. You declare them.
\item You do not hide your covenant. You wear it.
\item You may struggle with mercy. The flame judges; you must learn when to temper it.
\item You may be called to witness oaths outside of battle—marriages, treaties, adoptions. Your word carries weight.
\item Your scars are visible. Each one is a promise kept. Each one is a conversation starter—or a warning.
\end{itemize}

\subsection{Common Quirks}

\begin{dice}
\item You touch your gorget before every fight. It is not a nervous habit. It is a reminder.
\item You cannot stand to see a flame extinguished without cause. You will relight it—or explain why not.
\item You have a habit of repeating your vows under your breath when stressed. Others think you are praying. You are reminding yourself.
\item You refuse to fight in darkness without a light source. The flame must witness.
\end{dice}

\subsection{Sample NPC: Knight-Commander Dara of the Dawn Vigil}

\textbf{Archetype:} Dawn Warrior

\textbf{Patron:} Oath of Flame and Light

\textbf{Codex:} Engraved gorget, each vow a line of flame-light script

\textbf{Thiasos:} Ember, a spirit-hound of ash whose paw prints smolder for hours

\textbf{Personality:} Stern, weary, secretly uncertain. She has kept every vow she ever swore. She is not sure she remembers why.

\textbf{Conflict:} A vow she swore twenty years ago—to protect a village that no longer exists—has no witness but herself. Does the oath still bind? The flame says yes. Her heart is not sure.

\textbf{Quote:} \textit{"I swore an oath in fire. The fire does not forget. Neither do I. But the village is gone. The people are dust. Who am I protecting now?"}

\subsection{GM Guidance: Dawn Warrior Story Hooks}

\begin{itemize}
\item A vow the Dawn Warrior swore years ago resurfaces. A witness they had forgotten appears—and demands fulfillment.
\item An enemy begs for mercy. The flame demands judgment. The Dawn Warrior must choose.
\item A dawnfire Rite leaves a permanent mark: a scar that glows faintly in darkness, visible to all who know what it means.
\item The Dawn Warrior's flame gutters at a crucial moment—not because of magic, but because of doubt. The Patron is watching.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
