\section{Bloodsworn (The High Mother)}

\subsection{The Story of Sister Alva}

\epigraph{The soup is free. The price never is.}{From \textit{The Bitter Root}}

Sister Alva was not always a Bloodsworn. She was not always called Alva. Once, she was a midwife in a village that no longer remembers her name, serving a Cord that demanded payment for every mercy.

She attended a birth. Twins. The boy lived. The girl was stillborn. The mother was dying. The Cord demanded its price: a memory, a name, a last breath.

But the mother had held Alva's hand when no one else would. She had learned the herb craft. She had a gift for knots. She had asked for nothing.

So Alva cut the thread—not to collect, but to let go.

The price was terrible.

\subsection{The Scar That Weeps}

Alva's body is a ledger. Her back is covered in scars—each one a knot she has carried for someone else. Her arms are wrapped in tattoos that weep black ink when a debt comes due. Her wrists are bound in grey cord that tightens when she is tempted to speak a secret.

She has not spoken her own name in forty years. She is not sure she remembers it.

But the Hollowed remembers. The Hollowed is a child with milk-white eyes, its heart replaced with a bundle of grey thread. It sits in the corner of every room. It never blinks. It witnesses everything.

\subsection{The Knot That Held}

At a deathbed, a woman begged Alva to carry a secret—a crime her son had committed, a theft that would ruin him if revealed. Alva tied a knot in her grey cord. She carried the secret for thirty years. The knot tightened every night. She never untied it.

When the son lay dying, Alva sat beside him. She did not tell him the secret. She did not need to. He already knew.

\textit{"You carried it,"} he whispered. \textit{"For all those years."}

\textit{"The soup is free,"} Alva said. \textit{"The price never is. But some prices are paid by the one who carries, not the one who owes."}

She cut the knot. The cord loosened. The son died smiling.

\subsection{The Thread That Was Never Tied}

Alva was sent to collect a debt from a family that had not paid the Cord in three generations. The debt was not in coin—it was in memories. The family had forgotten the name of the grandmother who had first borrowed from the Cord.

Alva did not collect. She sat at their hearth. She drank their soup. She listened to their stories. And she tied a new knot—not to bind them, but to remind them.

\textit{"The Cord does not forget,"} she said. \textit{"Neither should you. But you do not owe me. You owe her. Remember her name, and the debt is paid."}

They remembered. The knot loosened. Alva left without collecting a single coin.

The Cord was not pleased. The High Mother was silent for a year. But Alva did not cut another knot. She simply waited. The Cord is patient. So is she.

\subsection{What the Bloodsworn Carries}

A Bloodsworn carries the debts of others. They carry secrets that would break lesser souls. They carry the weight of every mercy they have given and every mercy they have refused.

But most of all, they carry the knowledge that some threads are not meant to be cut. Some knots are not meant to be tied. And some debts are paid not by the debtor, but by the one who chooses to carry.

\vspace{1em}
\begin{center}
\textit{The thread held. The water carried. That is not a Rite. That is grace.}
\end{center}

\section{Bloodsworn (The High Mother)}

\subsection{Concept}

The Bloodsworn is a Daughter of the Unbroken Cord who has taken the covenant into battle. They are not healers who avoid conflict—they are warriors who understand that mercy has a ledger and that every wound is a debt. Their Codex is a set of flesh-tattoos, each Rite woven into their skin. Their Thiasos is a Hollowed child or heartless bird—a silent witness that never blinks. Their Rites store Fatigue as visible knots, appearing uninjured while their body weeps blood-thread.

A Bloodsworn's covenant is written on their flesh. Every knot is a promise kept. Every scar is a payment made. They do not hide what they owe.

\subsection{Core Attributes}

\begin{itemize}
\item \textbf{Primary:} Body, Spirit
\item \textbf{Secondary:} Wits
\item \textbf{Tertiary:} Presence
\end{itemize}

\subsection{Skills}

\begin{itemize}
\item \textbf{Core:} Melee, Medicine, Subterfuge
\item \textbf{Secondary:} Lore (Thresholds), Stealth
\item \textbf{Tertiary:} Sway, Insight
\end{itemize}

\subsection{Codex}

The Bloodsworn's Codex is woven into their flesh:

\begin{itemize}
\item \textbf{Flesh-Tattoos:} Rites tattooed onto the skin using ink mixed with ash from former Vessels and blood-milk from the Cord. The tattoos move—slowly, subtly—as debts accrue.
\item \textbf{Grey Cord:} A length of undyed wool tied around the wrist. Each knot is a promise kept. The cord tightens when the Bloodsworn is tempted to break faith.
\item \textbf{Scar-Plate:} A set of ritual scars across the back or chest. Each scar is a Rite invoked. The scars never fully heal.
\end{itemize}

\subsection{Thiasos}

\begin{itemize}
\item \textbf{Hollowed Child:} A child taken from a home that no longer remembers them, hollowed out by the High Mother's needle. Heart replaced with a bundle of grey thread. Eyes sewn open with silver wire. It witnesses. It never blinks. Cap 2, cannot be harmed by mundane means.
\item \textbf{Heartless Bird:} A small grey bird with milk-white eyes. It cannot sing. It watches. It weeps blood when a debt comes due. Cap 2, cannot be harmed by mundane means.
\item \textbf{The Scarlet Needle (Sentient):} A silver needle with a red thread already threaded. Speaks in whispers. Can stitch wounds closed—or stitch lips shut. (See Optional Rule: Sentient Thiasos)
\end{itemize}

\subsection{Signature Rites}

\subsubsection{Low Rite: Rite of the Flesh-Knot (4 XP)}

Store 1 Fatigue as a visible knot in your skin (a fresh scar, a weeping tattoo, a tightened cord). You may release the knot later to ignore 1 Fatigue from exertion. You may have up to Body rating knots at once.

\subsubsection{Standard Rite: Rite of the Silent Tithe (7 XP)}

Once per season, you may declare that a household owes the Cord a tithe—a memory, a shadow, or a first breath of a newborn. The tithe is collected within the week, often without the household's awareness. You gain 1 permanent segment of Influence within that household's community.

\subsubsection{High Rite: Rite of the Hollowed Union (12 XP)}

Join two bloodlines through a debt that will be paid by a child not yet conceived. Write the terms in blood on a strip of linen, then feed the linen to your Hollowed. The Hollowed's eyes film over for a day. One year later, the debt will manifest as you decreed.

\subsection{Signature Talents}

\subsubsection{Cord of Flesh (8 XP)}

Your body stores memories and debts. When you convert Harm → Fatigue (Monk ability), you may store up to Body rating points of Fatigue as Chained Knots, appearing uninjured (as fresh scars, weeping tattoos, or coiled threads beneath the skin). You have until the sunset after the next to perform a sacrifice or take 2 Harm.

\subsubsection{The Open Palm, Closed Cord (8 XP)}

Your unarmed strikes or subtle touch carry Widow's Measure. Instead of dealing Harm, you may mark a Promise Clock [4] on the target—they owe the Cord a truth or grief. If they already owe the coven, you may activate that debt: they suffer Position -1 for the scene, or you gain Effect +1 against them.

\subsubsection{The Red Thread (4 XP)}

Once per scene, you may tie a small knot in a cord, thread, or cloth within reach. For the rest of the scene, you and allies within Near range gain +1 die to a single type of roll (choose when you tie the knot: stealth, persuasion, defense, or navigation). The knot must remain tied; if undone, the benefit ends.

\subsection{Sample Build — Tier I (30-40 XP)}

\begin{itemize}
\item \textbf{Attributes:} Body 3 (9 XP), Spirit 2 (6 XP), Wits 2 (6 XP), Presence 1 (3 XP) — Total 24 XP
\item \textbf{Skills:} Melee 1 (2 XP), Medicine 1 (2 XP), Subterfuge 1 (2 XP), Stealth 1 (2 XP) — Total 8 XP
\item \textbf{Talents:} Flesh Codex (2 XP), The Red Thread (4 XP), Scarred Resilience (3 XP) — Total 9 XP
\item \textbf{Remaining XP:} 0 (30 XP base + bonds/complications as needed)
\end{itemize}

\subsection{Sample Build — Tier III (91-150 XP)}

\begin{itemize}
\item \textbf{Attributes:} Body 4 (18 XP), Spirit 3 (12 XP), Wits 3 (12 XP), Presence 2 (6 XP) — Total 48 XP
\item \textbf{Skills:} Melee 2 (4 XP), Medicine 2 (4 XP), Subterfuge 2 (4 XP), Lore (Thresholds) 2 (6 XP), Stealth 2 (4 XP), Sway 1 (2 XP) — Total 24 XP
\item \textbf{Talents:} Flesh Codex, The Red Thread, Scarred Resilience, Cord of Flesh (8 XP), The Open Palm, Closed Cord (8 XP) — Total 18 XP (cumulative)
\item \textbf{Rites:} Rite of the Flesh-Knot (4 XP), Rite of the Silent Tithe (7 XP) — Total 11 XP (cumulative)
\item \textbf{Total XP:} 48 + 24 + 18 + 11 = 101 XP
\end{itemize}

\subsection{Roleplaying a Bloodsworn}

\begin{itemize}
\item You do not give gifts. You extend credit. The soup is free. The price never is.
\item You may be called to witness deathbed confessions, to tend the sick, or to collect debts that have compounded for generations.
\item You may struggle with the weight of what you carry. Every knot is a debt. Every scar is a payment. The Cord does not forget. Neither should you.
\item You may be asked to cut a thread—to release a debt, to forgive a promise, to let someone go free. The Cord does not forbid mercy. It asks only that you remember.
\item Your Hollowed is always watching. You have stopped noticing. Others never do.
\end{itemize}

\subsection{Common Quirks}

\begin{dice}
\item You tie and untie knots in a length of grey cord when thinking. It is not nervous energy. It is accounting.
\item You sometimes touch your tattoos as if checking a pulse. The tattoos are warm when a debt is near.
\item You cannot accept a gift without naming its price aloud. Even a cup of tea. Even a compliment.
\item Your Hollowed sits in the corner of every room. You have never introduced it. No one has asked.
\end{dice}

\subsection{Sample NPC: Sister Alva of the Unbroken Cord}

\textbf{Archetype:} Bloodsworn

\textbf{Patron:} The High Mother

\textbf{Codex:} Flesh-tattoos that weep when debts come due

\textbf{Thiasos:} A Hollowed child with milk-white eyes that never blink

\textbf{Personality:} Quiet, steady, inexhaustible. She has witnessed a thousand confessions. She has never repeated one.

\textbf{Conflict:} A debt she collected thirty years ago has come due again—not from the original debtor, but from their grandchild. The Cord does not distinguish between generations. Alva must decide: collect, or cut the thread for the first time.

\textbf{Quote:} \textit{"The soup is free. The price never is. They ask me why I don't eat. I tell them I am not hungry. I do not tell them I am already full—of others' debts."}

\subsection{GM Guidance: Bloodsworn Story Hooks}

\begin{itemize}
\item A Hollowed's gaze lingers on a character. They feel watched. Next roll involving secrecy is Desperate.
\item A thread of grey cord appears tied around a character's wrist. They cannot recall when it happened.
\item The High Mother's attention falls on a character. They dream of the Cord. When they wake, they have forgotten a significant memory (GM chooses—a face, a secret, a location).
\item A Bloodsworn is asked to cut a thread that has held for generations. The Cord will remember. The question is: will they?
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
