\section{Ghe'hai Operative (Lethai Syncretic)}

\subsection{The Story of Silk-Weaver Mira}

\epigraph{She wore a mask of woven silk. Beneath it, she was anyone—and no one.}{From the Archives of the Silk Vigil}

Mira was not born into the Ghe'hai. She was recruited—or perhaps stolen is the better word. A child of the Crimson Valley, she showed an early gift for knots and silences. The Lethai do not waste such gifts.

She trained for ten years in the ways of the Silk Vigil: how to read a room by the tension of its threads, how to kill without drawing blood, how to negotiate from a position of absolute weakness. Her teachers were masks, not faces. She never knew their names.

Her Codex is a woven silk mask, each thread a memory of a face she has worn. Her Thiasos is a spider-spirit that weaves threads between people. Her Rites are the knots she ties and unties with a whisper.

She has killed nine people. She has saved twice that many. She does not remember the difference.

\subsection{The Mask That Fits}

Mira carries seven masks in a silk bag. Each is a different face—a merchant, a servant, a noble, a beggar. She has worn them all. She is not sure which one is hers.

The masks were not carved. They were grown—from mulberry trees, watered with tears and memory. When she puts one on, her face becomes the mask's. Her voice becomes the mask's. Her soul becomes the mask's—for a time.

She never wears the same mask twice in the same city.

\subsection{The Thread That Binds}

In the Crimson Valley, the Ghe'hai do not fight. They negotiate. Their weapons are silk cords and silvered tongues. A Ghe'hai operative can end a feud with a sentence, start a war with a whisper, or simply vanish into the crowd like smoke through a keyhole.

Mira's signature is a single strand of red silk, tied around a door handle or a bedpost or a throat. It means: \textit{I was here. You owe me.}

No one knows what the debt is until she collects.

\subsection{The River at Midnight}

Mira was sent to retrieve a locket from a dead woman's estate. The locket contained nothing—only a lock of auburn hair and a scrap of red thread. Her client would not say why it mattered.

She found the locket in a vault beneath a burned chapel. The vault was guarded by a ghost—a woman with a brass hook and no shadow. The ghost did not attack. She simply asked: \textit{Why are you here?}

Mira answered: \textit{To deliver what was never stolen.}

The ghost stepped aside. The locket was warm. The thread was red. Mira left it on a riverbank at dawn. She did not look back.

\subsection{What the Ghe'hai Operative Carries}

A Ghe'hai operative carries silk and steel. They carry masks and memories. They carry the weight of every face they have worn, every debt they have collected, every secret they have kept.

But most of all, they carry the knowledge that the strongest weapon is not a blade—it is a thread, tied at the right moment, witnessed by no one.

\vspace{1em}
\begin{center}
\textit{The thread held. The water carried. That is not a Rite. That is grace.}
\end{center}

\section{Ghe'hai Operative (Lethai Syncretic)}

\subsection{Concept}

The Ghe'hai operative is an elite warrior-assassin-negotiator of the Lethai courts. They are not soldiers. They are not spies. They are something in between—a diplomat who carries a knife, an assassin who prefers not to use it, a whisper who can become a scream.

Their Codex is a silk mask or a spider-silk cord. Their Thiasos is a spider-spirit that weaves threads between people, or a sentient blade wrapped in silk that never dulls. Their Rites are subtle, social, and often invisible—a veil, a binding thread, a whispered command.

A Ghe'hai operative does not fight fair. They do not fight at all if they can avoid it. But when they do, the enemy is already tangled.

\subsection{Core Attributes}

\begin{itemize}
\item \textbf{Primary:} Presence, Wits
\item \textbf{Secondary:} Body
\item \textbf{Tertiary:} Spirit
\end{itemize}

\subsection{Skills}

\begin{itemize}
\item \textbf{Core:} Deception, Insight, Stealth
\item \textbf{Secondary:} Melee, Sway, Subterfuge
\item \textbf{Tertiary:} Lore (Lethai customs), Performance
\end{itemize}

\subsection{Codex}

The Ghe'hai operative's Codex is a tool of subtlety and disguise:

\begin{itemize}
\item \textbf{Silk Mask:} A mask of woven silk that shifts patterns. Rites are woven into the silk, invisible until touched. Wearing it grants +1 die to Deception.
\item \textbf{Spider-Silk Cord:} A length of pale, almost invisible silk. Stronger than steel. Each knot is a Rite waiting to be invoked.
\item \textbf{Woven Glove:} A silk glove on the off-hand. It can tie threads, bind lips, or simply rest—waiting.
\end{itemize}

\subsection{Thiasos}

\begin{itemize}
\item \textbf{Spider-Spirit:} A translucent spider the size of a hand. It weaves threads between people, binding them in webs of obligation. Can sense hidden relationships (debt, affection, rivalry) within Near range. Cap 2, cannot be harmed by mundane means.
\item \textbf{Memory-Echo:} A spirit that cannot speak, but can replay sounds it has witnessed—a confession, a scream, a name spoken in the dark. Cap 2, Harm 3.
\item \textbf{The Silk Mask (Sentient):} A mask that speaks in a neutral, androgynous voice. Can assume the appearance of any face the Runekeeper has touched (once per session, for one scene). (See Optional Rule: Sentient Thiasos)
\end{itemize}

\subsection{Signature Rites}

\subsubsection{Low Rite: Rite of the Silk Binding (4 XP)}

\textbf{Tags:} [BIND], [SILENCE], [THRESHOLD]

Immobilize a target with invisible threads. They suffer -1 die to all physical actions for one scene. Breaking free requires a Body + Athletics test (DV 4) and an action.

\subsubsection{Low Rite: Rite of the Velvet Whisper (4 XP)}

\textbf{Tags:} [SILENCE], [MANIPULATION], [BOOST]

Your private words carry irresistible charm. Gain +1 die to Sway a single target in confidential conversation. On a success, they may also reveal a minor secret without realizing it.

\subsubsection{Standard Rite: Rite of the Mask of Many Faces (7 XP)}

\textbf{Tags:} [TRANSFORM], [VEIL], [DECEPTION]

Assume the appearance of any person you have touched for one scene. Your voice, build, and mannerisms change. You do not gain their memories or skills. Detecting the disguise requires a contested Insight roll (DV 4).

\subsubsection{Standard Rite: Rite of the Spinner's Web (8 XP)}

\textbf{Tags:} [BOND], [FATE], [MANIPULATION]

Link two actors with an invisible strand. For the rest of the scene, whenever one acts, the other is dragged into consequence. You may trigger this linkage once to force 1 SB on the linked target after they perform a significant action.

\subsubsection{High Rite: Rite of the Unspoken Question (12 XP)}

\textbf{Tags:} [MIND], [TRUTH], [FEAR]

Force a target within Near range to answer one question truthfully. They may resist with Resolve (DV 5). On a failure, they answer—one sentence, no evasion. On a success, they suffer 1 Fatigue from the effort of resistance.

\subsection{Signature Talents}

\subsubsection{Ghe'hai's Escape (6 XP)}

\textbf{Requirement:} Lethai Syncretic, Wits 3+

Once per session, when you are cornered or capture is inevitable, you may declare that you prepared an escape route. You vanish, reappearing in a safe location of the GM's choosing, but you leave behind a token (a mask, a silk cord, a single word) that can be used to track you.

\subsubsection{Mask of Many Faces (4 XP)}

\textbf{Requirement:} Lethai Syncretic

You can change your appearance as an action. This is not illusion—it is physical transformation. You may alter your face, build, and voice. You cannot mimic a specific person without a token (hair, signature, worn clothing). The change lasts until you choose to revert.

\subsubsection{The Unseen Thread (8 XP)}

\textbf{Requirement:} Lethai Syncretic, Presence 3+

Once per session, you may declare that a target owes you a debt from a previous, forgotten interaction. They must test Spirit + Resolve (DV 4) or accept the truth of the debt. If they fail, they owe you one significant service or truth.

\subsection{Sample Build — Tier I (30-40 XP)}

\begin{itemize}
\item \textbf{Attributes:} Presence 3 (9 XP), Wits 3 (9 XP), Body 2 (6 XP), Spirit 1 (3 XP) — Total 27 XP
\item \textbf{Skills:} Deception 2 (4 XP), Insight 1 (2 XP), Stealth 1 (2 XP) — Total 8 XP
\item \textbf{Talents:} Flesh Codex (2 XP), Ghe'hai's Escape (6 XP), Mask of Many Faces (4 XP) — Total 12 XP (using bonds/complications for remaining XP)
\item \textbf{Remaining XP:} -3 (use 1 bond for +2 XP, 1 complication for +4 XP to reach 41 XP)
\end{itemize}

\subsection{Sample Build — Tier III (91-150 XP)}

\begin{itemize}
\item \textbf{Attributes:} Presence 4 (15 XP), Wits 4 (15 XP), Body 3 (9 XP), Spirit 2 (6 XP) — Total 45 XP
\item \textbf{Skills:} Deception 3 (9 XP), Insight 2 (6 XP), Stealth 2 (6 XP), Melee 2 (6 XP), Sway 2 (6 XP), Subterfuge 1 (2 XP) — Total 35 XP
\item \textbf{Talents:} Flesh Codex, Ghe'hai's Escape, Mask of Many Faces, The Unseen Thread (8 XP), Engraved Defense (6 XP) — Total 20 XP (cumulative)
\item \textbf{Rites:} Rite of the Silk Binding (4 XP), Rite of the Velvet Whisper (4 XP), Rite of the Mask of Many Faces (7 XP) — Total 15 XP (cumulative)
\item \textbf{Total XP:} 45 + 35 + 20 + 15 = 115 XP
\end{itemize}

\subsection{Roleplaying a Ghe'hai Operative}

\begin{itemize}
\item You do not fight fair. You fight smart. The enemy does not need to die—only to be neutralized.
\item You may be called to negotiate treaties, to infiltrate hostile courts, or to retrieve secrets that cannot be stolen by force.
\item You may struggle with identity. The masks you wear are not just disguises—they are selves. You may forget which one is yours.
\item You may be asked to kill. The Ghe'hai do not forbid it. But they prefer a living asset to a dead one. You must choose when the knife is necessary.
\item Your webs are everywhere. Every debt you collect, every thread you tie, is a resource—or a liability.
\end{itemize}

\subsection{Common Quirks}

\begin{dice}
\item You touch your mask before speaking to a stranger. It is not a nervous habit. It is a choice.
\item You cannot stand to leave a thread untied. A loose end is a vulnerability.
\item You have a habit of speaking in conditional tense—"If you were to..." "One might consider..." It is not indecision. It is leverage.
\item You refuse to make eye contact for more than a heartbeat. The eyes are the first thread.
\end{dice}

\subsection{Sample NPC: Mask-Weaver Sera of the Silk Vigil}

\textbf{Archetype:} Ghe'hai Operative

\textbf{Patron:} Lethai Syncretic (Inaea / Isoka / The Witness)

\textbf{Codex:} A silk mask that shifts patterns

\textbf{Thiasos:} A spider-spirit that weaves threads between people

\textbf{Personality:} Calm, watchful, unreadable. She has worn a hundred faces. She is not sure which one is hers.

\textbf{Conflict:} A thread she tied twenty years ago—between a mother and a daughter—has tightened. The daughter is dying. The mother is far away. The thread remembers. Sera must decide: cut it, or let it hold until the end.

\textbf{Quote:} \textit{"I am not the one who tied that thread. I only witnessed it. But the witness is also the weaver. And the weaver is always responsible."}

\subsection{GM Guidance: Ghe'hai Operative Story Hooks}

\begin{itemize}
\item A mask the Ghe'hai operative wore in another city has been recognized. Someone from that past life is here—and they are not friendly.
\item A thread the operative tied years ago has tightened at the worst possible moment. A debt is coming due.
\item The operative is asked to kill someone who does not deserve to die. The Lethai code says: negotiate first. The client says: no negotiation.
\item The operative's silk mask begins to speak in a voice they do not recognize. It is asking for a name. The operative does not know which name to give.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.