\section{Beast Warrior (The Savage Heart)}

\subsection{The Story of Mkara the Leopard's Shadow}

\epigraph{She wore the pelt of the beast that killed her father. Now she hunts in its skin.}{From the Hunt-Songs of Sekogo}

Mkara was not always a Beast Warrior. Once, she was a child who watched a leopard drag her father into the tall grass. She was twelve. She had no weapon. She had no magic. She had only her hands and her teeth and a scream that no one heard.

She followed the blood trail for three days. On the third night, she found the leopard sleeping in a cave, its belly full of her father's flesh. She did not have a knife. She used a sharpened stone. It took a long time. The leopard did not die quietly.

She wore its pelt home. The village elders declared her \textit{bint'leopard}—the Leopard's Daughter. They taught her the old ways: how to move without sound, how to kill without waste, how to offer the first bite to the earth.

Her Codex is the leopard's tooth, worn on a leather thong. Her Thiasos is the leopard's spirit, which hunts beside her still. Her Rites are the hunts she undertakes—each one a sacrament, each kill a prayer.

\subsection{The Scar That Remembers}

Mkara has a scar across her ribs—three parallel lines, claw marks from the leopard's death spasm. She does not hide it. She oils it. It is the first scar she earned, but not the last.

Each scar on her body is a hunt remembered. She can touch a scar and recall the scent of the prey, the weight of the strike, the taste of the earth at the moment of the kill. The Savage Heart does not let her forget.

\subsection{The Hunt That Never Ends}

The leopard's spirit hunts beside Mkara. It is not a pet. It is not a servant. It is a pack-brother. It speaks in growls and silences. It does not ask permission. It simply stalks, and Mkara follows.

They have hunted together for twenty years. They have never lost a trail. They have never failed to make the kill. They have never spoken a word to each other—not one. They do not need to.

The hunt is the conversation. The kill is the punctuation.

\subsection{The River at Dawn}

Mkara was hired to hunt a murderer who had fled into the Mistlands. The trail led to the Black River. The murderer was already dead—killed by someone else, someone with a brass hook and no shadow.

Mkara did not mourn the lost quarry. She knelt at the riverbank and offered a piece of dried meat to the water. The leopard's spirit drank beside her.

She did not know whose ghost had stolen her kill. She did not care. The hunt was over. The Savage Heart was satisfied.

\subsection{What the Beast Warrior Carries}

A Beast Warrior carries the pelt of their first kill. They carry the tooth of the beast that changed them. They carry the scars of every hunt that nearly killed them.

But most of all, they carry the knowledge that they are not the apex predator. The wild is patient. The wild is hungry. And the wild is always watching.

\vspace{1em}
\begin{center}
\textit{The thread held. The water carried. That is not a Rite. That is grace.}
\end{center}

\section{Beast Warrior (The Savage Heart)}

\subsection{Concept}

The Beast Warrior is a predator who hunts as sacrament. They are not guardians of the wild or stewards of nature's balance. They are its teeth and claws. Power comes from taking, from proving strength, from the raw act of predation.

Their Codex is a predator's tooth or a pelt worn as a cloak. Their Thiasos is a wolf-ancestor, a great cat's spirit, or a sentient blade carved from the fang of a legendary beast. Their Rites are the hunts they undertake—each one a prayer, each kill a sacrament.

A Beast Warrior does not fight fair. They fight to win. The hunt does not care about honor. It cares about the kill.

\subsection{Core Attributes}

\begin{itemize}
\item \textbf{Primary:} Body, Wits
\item \textbf{Secondary:} Spirit
\item \textbf{Tertiary:} Presence
\end{itemize}

\subsection{Skills}

\begin{itemize}
\item \textbf{Core:} Survival, Stealth, Melee
\item \textbf{Secondary:} Athletics, Investigation (tracking), Intimidation
\item \textbf{Tertiary:} Lore (local wildlife), Endurance
\end{itemize}

\subsection{Codex}

The Beast Warrior's Codex is a trophy of the hunt:

\begin{itemize}
\item \textbf{Predator's Tooth:} A large canine tooth on a leather thong. It is warm when prey is near. It grows heavier with each kill.
\item \textbf{Pelt Cloak:} The cured hide of an apex predator (bear, wolf, lion, tiger, leopard). Wearing it grants +1 die to Intimidation and +1 Armor against cold.
\item \textbf{Claw Necklace:} A string of claws from worthy kills. Each claw holds a memory of the hunt. The necklace clicks when the hunt begins.
\end{itemize}

\subsection{Thiasos}

\begin{itemize}
\item \textbf{Wolf-Ancestor:} A grey wolf the size of a pony, with eyes that reflect the Runekeeper's first kill. Tracks across any terrain, fights beside its pack-brother. Cap 3, Harm 5.
\item \textbf{Great Cat Spirit:} A translucent leopard or tiger that moves without sound. Can stalk a target for hours without being seen. Cap 3, Harm 4.
\item \textbf{The Fang (Sentient):} A blade carved from the tooth of a legendary beast. It speaks in growls. It hungers. It demands to be fed. (See Optional Rule: Sentient Thiasos)
\end{itemize}

\subsection{Signature Rites}

\subsubsection{Low Rite: Rite of the Predator's Gaze (4 XP)}

\textbf{Tags:} [MARK], [FEAR], [TRACK]

Mark a single target as your prey. Gain +1 die on all rolls to track, pursue, or intimidate that target. The effect lasts until the target escapes or is killed.

\subsubsection{Low Rite: Rite of the Blood Scent (5 XP)}

\textbf{Tags:} [SENSE], [FEAR], [TRACK]

You scent the air like a beast. Gain +2 dice to Investigation or Survival when tracking by smell. You may also detect fear—you sense if a creature within Near is Frightened.

\subsubsection{Standard Rite: Rite of the Shifting Shape (8 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST]

Assume the form of a predator for the scene. Choose bear (tank form: +1 Armor, +2 dice to Body rolls, movement reduced) or wolf (striker form: +2 dice to Agility and Stealth, +1 band movement). You may speak but cannot use tools or cast other Rites while transformed.

\subsubsection{Standard Rite: Rite of the Pack Hunt (7 XP)}

\textbf{Tags:} [TEAMWORK], [BEAST], [BOOST]

You and up to three willing allies form a hunting pack. While within Near of each other, all gain +1 die to attacks against a single designated prey, and the first ally to attack each round gains +1 Effect.

\subsubsection{High Rite: Rite of the Savage Transformation (13 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST], [FEAR]

Fully transform into an apex predator for one scene. You gain the physical attributes of the beast: Body +2, +2 Armor from thick hide, and your melee attacks deal +2 Harm. You cannot speak or use tools. You may end the transformation early as an action. When the transformation ends, you mark 2 Fatigue.

\subsection{Signature Talents}

\subsubsection{Beast Form (6 XP)}

\textbf{Requirement:} The Savage Heart, Body 3+

Once per session, you may partially transform into a predator for one scene. Choose one: +2 dice to Stealth (leopard/tiger), +2 dice to Intimidation (bear/lion), or +1 band to movement (wolf/hound). You cannot use this talent while already transformed.

\subsubsection{Blood Hunt (4 XP)}

\textbf{Requirement:} The Savage Heart

When you are tracking a target that you have wounded, you gain +1 die to all tracking and pursuit rolls. The scent of their blood is a trail you cannot lose.

\subsubsection{Alpha's Roar (8 XP)}

\textbf{Requirement:} The Savage Heart, Presence 3+

Once per session, you may make a terrifying sound—a roar, a howl, a scream. All enemies within Near range must test Resolve (DV 4) or become Shaken (-1 die to all actions) for one exchange. Does not affect creatures that are immune to fear.

\subsection{Sample Build — Tier I (30-40 XP)}

\begin{itemize}
\item \textbf{Attributes:} Body 3 (9 XP), Wits 3 (9 XP), Spirit 2 (6 XP), Presence 1 (3 XP) — Total 27 XP
\item \textbf{Skills:} Survival 2 (4 XP), Stealth 1 (2 XP), Melee 1 (2 XP) — Total 8 XP
\item \textbf{Talents:} Flesh Codex (2 XP), Blood Hunt (4 XP), Scarred Resilience (3 XP) — Total 9 XP
\item \textbf{Remaining XP:} -4 (use 2 bonds for +4 XP to reach 40 XP)
\end{itemize}

\subsection{Sample Build — Tier III (91-150 XP)}

\begin{itemize}
\item \textbf{Attributes:} Body 4 (18 XP), Wits 4 (15 XP), Spirit 3 (12 XP), Presence 2 (6 XP) — Total 51 XP
\item \textbf{Skills:} Survival 3 (9 XP), Stealth 3 (9 XP), Melee 3 (9 XP), Athletics 2 (6 XP), Intimidation 2 (6 XP), Investigation (tracking) 1 (2 XP) — Total 41 XP
\item \textbf{Talents:} Flesh Codex, Blood Hunt, Scarred Resilience, Beast Form (6 XP), Alpha's Roar (8 XP) — Total 20 XP (cumulative)
\item \textbf{Rites:} Rite of the Predator's Gaze (4 XP), Rite of the Blood Scent (5 XP), Rite of the Shifting Shape (8 XP) — Total 17 XP (cumulative)
\item \textbf{Total XP:} 51 + 41 + 20 + 17 = 129 XP
\end{itemize}

\subsection{Roleplaying a Beast Warrior}

\begin{itemize}
\item You do not fight for glory. You hunt for survival. The kill is sacrament; the meat is blessing.
\item You may be called to track fugitives, hunt dangerous beasts, or lead a pack in war.
\item You may struggle with the hunger. The Savage Heart does not ask you to be gentle. It asks you to be true—to the hunt, to the pack, to yourself.
\item You may be asked to kill something that does not deserve to die. The Heart does not care about deserving. The Heart is hungry.
\item Your scars are your resume. Each one is a story. Each one is a warning.
\end{itemize}

\subsection{Common Quirks}

\begin{dice}
\item You sniff the air when you enter a room. You are not being rude. You are scenting for threats.
\item You cannot sleep indoors. The walls are a cage. You will find a way to sleep under the sky, even if it is just a courtyard.
\item You have a habit of offering the first bite of any meal to the earth. It is not prayer. It is tribute.
\item You growl when angry. You do not notice you are doing it. Others notice.
\end{dice}

\subsection{Sample NPC: Mkara of the Leopard's Shadow}

\textbf{Archetype:} Beast Warrior

\textbf{Patron:} The Savage Heart

\textbf{Codex:} A leopard's tooth on a leather thong

\textbf{Thiasos:} A leopard's spirit that hunts beside her

\textbf{Personality:} Silent, patient, deadly. She has never lost a hunt. She has never needed to speak.

\textbf{Conflict:} A child has wandered into the territory of a wounded leopard. The leopard is dangerous. The child is innocent. The hunt says: kill. The heart says: wait. The leopard is watching.

\textbf{Quote:} \textit{"The leopard does not hate you. It does not love you. It is hungry. That is all. That is enough."}

\subsection{GM Guidance: Beast Warrior Story Hooks}

\begin{itemize}
\item The Beast Warrior's prey escapes—not because of poor skill, but because the prey was not worthy. The Savage Heart is displeased.
\item A rival predator enters the Beast Warrior's territory. It does not attack. It watches. It is judging.
\item The Beast Warrior's pelt cloak begins to itch. The beast whose skin it was is not entirely dead. It dreams.
\item The hunt lasts too long. The Beast Warrior marks 1 Fatigue. The hunger grows. The Savage Heart whispers: \textit{faster}.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
