% Threadweaver's Spellbook
% A Fate's Edge Expansion
% Compiled by Lysandra of the Amber Gate

\documentclass[11pt]{book}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{hyperref}

\definecolor{thepyrgos}{RGB}{120, 60, 30}
\definecolor{danger}{RGB}{180, 40, 40}

\titleformat{\chapter}{\normalfontℎuge\bfseries\color{thepyrgos}}{þechapter}{20pt}{}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\slshape Threadweaver's Spellbook}
\fancyhead[RE,LO]{\slshape ≤ftmark}
\fancyfoot[C]{þepage}
\renewcommand{ℎeadrulewidth}{0.4pt}

ℎypersetup{
    colorlinks=true,
    linkcolor=thepyrgos,
    urlcolor=blue,
}

\begin{document}

\frontmatter
\title{Threadweaver's Spellbook}
\author{Lysandra of the Amber Gate\\\small Master of Spellcraft, Thepyrgos}
\date{Under the seal of the Synod, Third Tier\\Year of the Broken Crown}
\maketitle

\tableofcontents

\chapter*{Preface: A Letter to the Inquisitive}
\addcontentsline{toc}{chapter}{Preface: A Letter to the Inquisitive}

You have opened a book that should not exist. The Synod licenses texts on practical metaphysics; they do not license grimoires that teach the unsanctioned to set fires with a whisper. Yet here you are, and here I am, and the Weave is patient.

I am Lysandra of the Amber Gate. For thirty years I have taught Spellcraft at the University of Thepyrgos, and for thirty years I have watched students mispronounce [HEAL] and set their own sleeves ablaze. This book is for the ones who survived. It is also for the ones who will not, because they will ignore every warning I write and leap toward danger with a grin. I salute those students. They are the ones who discover new tags.

What follows is a grammar of the raw Weave. You will find no incantations to memorise, no patron to beseech, no symbol to polish. You will find intention, tag, and consequence. The rest is practice.

I have divided the work into eight chapters. The first teaches the language of tags. The next four catalogue spells by their purpose: humble cantrips, defensive arts, the unseen hand, and the elemental chorus. The sixth chapter is the one the Synod would burn: the dangerous door, where reality folds. The seventh chapter is a meditation on backlash−−−the Weave's receipt for your ambition. The eighth chapter is advice on practice, because theory without sweat is a lecture without a desk.

The appendices that follow contain talents, builds, a complete tag reference, a lexicon of Thepyrgos slang, a collection of instructive failures, and an examination scenario for those who wish to test their skill. The Caster's Ledger provides blank templates for recording your own spells.

If you are a student of the University, you may consider this a supplement to my lectures. If you are a hedge− mage without a patron, you may consider this a lifeline. If you are a Chain− Lantern inquisitor who has followed the paper trail to this page, you may consider it evidence. I will not hide.

The Weave does not care about your credentials. It only cares that you speak clearly.

\textit{−−− Lysandra of the Amber Gate}
\textit{Thepyrgos, in the hour before the eighth bell}

\mainmatter

\chapter{The Grammar of the Weave}
∈put{chapters/grammar.tex}

\chapter{Cantrips of the Humble Hand}
∈put{chapters/cantrips.tex}

\chapter{The Defensive Arts}
∈put{chapters/defensive.tex}

\chapter{The Unseen Hand}
∈put{chapters/unseen.tex}

\chapter{The Elemental Chorus}
∈put{chapters/elemental.tex}

\chapter{The Dangerous Door}
∈put{chapters/dangerous.tex}

\chapter{The Weave's Receipt}
∈put{chapters/backlash.tex}

\chapter{On Practice}
∈put{chapters/practice.tex}

\appendix

\chapter{Talents of the Threadweaver}
∈put{appendices/talents.tex}

\chapter{Builds for the Threadweaver}
∈put{appendices/builds.tex}

\chapter{Tags of the Weave}
∈put{appendices/tags.tex}

\chapter{Common Failures and Their Lessons}
∈put{appendices/complications.tex}

\chapter{A Short Lexicon of Thepyrgos Slang}
∈put{appendices/lexicon.tex}

\chapter{The Examination of the Amber Gate}
∈put{appendices/amber_gate.tex}

\chapter{The Caster's Ledger}
∈put{chapters/ledger.tex}

\backmatter
\chapter*{Colophon}
\addcontentsline{toc}{chapter}{Colophon}

This book was written on paper that had been a customs manifest, a rejected petition for readmission to the Synod's good graces, and a letter from a student who set fire to her own desk. The ink is lamp− black cut with the ash of a burnt syllabus and a single drop of my own blood—for luck, not for binding. The binding is the spine of a disgraced scholar's own dissertation. The margins contain marginalia from three of my students. I have not erased it. They earned the space.

The Weave does not forget. Neither do I.

If you have read this far, you have already begun. Light a candle. Name a tag. Roll the dice. The backlash is yours.

I have taught for forty years. I have seen students become masters, masters become legends, and legends become cautionary tales. I have no idea which of those categories I belong to. The Weave does not tell me. It only sends receipts.

One final piece of advice: when the Weave speaks, listen. When it laughs, laugh with it. When it demands payment, pay. And when it offers you a receipt, keep it. You will need it for your own ledger.

The road is south. The lamp is borrowed. The ink is fresh.

\textit{−−− Lysandra of the Amber Gate}
\textit{Thepyrgos, in the hour before the eighth bell}
\textit{The candle is still burning.}

\end{document}
