\section{Grimmir — The Wild Speaker}

\subsection{Domain and Epithets}

Grimmir is the Old Man of the Forest, the presence that watches from the shadow of the oldest oak, the cold breath on the back of the neck when the path grows dark. He is not kind, nor is he cruel—he is wild, and the wild does not negotiate. It demands, and it takes. His followers learn to read the language of blood on snow, the geometry of root and stone, and the weight of a heart offered at the turning of the year.

Those who walk his path become keepers of the old ways: druids who know when a grove must be fed, rangers who leave the first kill for the soil, and hedge-folk who whisper the names of spirits not to be spoken aloud. Their magic is not the gentle coaxing of garden plants but the raw, patient hunger of the forest floor—a hunger that swallows bones and grows mushrooms in the shape of the dead.

\subsection{Cultural Origins}

Grimmir appears wherever the wild presses against the works of civilization:

\begin{itemize}
\item In \textbf{Aeler}, he is the Root-Father, patron of deep fungus and stone-lichen. His shrines are carved into living rock where moss is never cleared.
\item In \textbf{Ykrul}, he is a sky-wolf who herds the seasons; shamans invoke Grimmir to predict the spring thaw.
\item In \textbf{Aelaerem}, he is the Thorn-Father, keeper of hedgerow magic. Villagers tie red thread around his roadside stones before planting.
\item In \textbf{Valewood}, he is the Old Oak, the spirit that watches from the heart of the forest. The Lethai leave offerings of bread and salt at his roots.
\end{itemize}

\subsection{Warrior Archetype: The Root-Knight}

A Runic Warrior of Grimmir is not a steward of gardens. They are a \textit{Root-Knight}—a wild warrior who fights with root and thorn, who wears the forest as armor and carries the season's hunger in their blade. Their Codex is a carved staff or living bark. Their Thiasos is a wolf-ancestor that hunts beside them, a fungal familiar that speaks in spores, or a sentient blade grown from a lightning-struck oak.

\subsection{Codex Form}

The Root-Knight's Codex is a living thing:

\begin{itemize}
\item \textbf{Carved Staff:} A staff of oak or ash, carved with seasonal runes. It grows leaves in spring and sheds them in autumn. It is never truly dead.
\item \textbf{Living Bark:} A strip of bark taken from a sacred tree. Rites are carved into it. It remains alive, weeping sap when a Rite is invoked.
\item \textbf{Thorn-Cord:} A braided cord of hawthorn and bramble. It draws blood when touched. The blood feeds the Rite.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Wolf-Ancestor:} A grey wolf the size of a pony, with eyes that reflect scenes from the Runekeeper's childhood. It tracks across any terrain and teaches forgotten battle-songs.
\item \textbf{Fungal Familiar:} A pale fungus that grows from a crack in the Runekeeper's armor. It speaks in spores and can detect decay, rot, or corruption within Near range.
\item \textbf{The Root-Staff (Sentient):} A walking staff carved from living wood. It speaks in the rustle of wind through branches. It can sense the health of nearby plants.
\end{itemize}

\subsection{Core Mechanic: The Seasonal Bond}

At the start of each season (spring, summer, autumn, winter), the Root-Knight gains a seasonal boon:

\begin{itemize}
\item \textbf{Spring:} +1 die to healing and growth-related rolls.
\item \textbf{Summer:} +1 die to strength and harvest-related rolls.
\item \textbf{Autumn:} +1 die to preservation and preparation-related rolls.
\item \textbf{Winter:} +1 die to endurance and hiding-related rolls.
\end{itemize}

The boon lasts until the next season. The Root-Knight may change their bond by performing a seasonal ritual (one day, requires a threshold and an offering).

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Blood Offering (4 XP)}

\textbf{Tags:} [SACRIFICE], [NATURE], [STEALTH]

\textbf{Threshold:} A drop of your own blood (1 Fatigue) or the blood of a freshly killed animal.

The forest accepts your offering. For the remainder of the scene, you gain +1 die to all Survival and Stealth rolls in natural terrain. You may ask the GM one yes/no question about a threat hidden in the wilderness.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Predator's Gaze (5 XP)}

\textbf{Tags:} [MARK], [FEAR], [TRACK]

\textbf{Threshold:} A tooth from a predator that killed its own prey.

Mark a single target as your prey. Gain +1 die on all rolls to track, pursue, or intimidate that target. The effect lasts until the target escapes or is killed.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shifting Shape (8 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST]

\textbf{Threshold:} A pelt or mask from the beast whose form you take.

Assume the form of a predator for the scene. Choose bear (tank form: +1 Armor, +2 dice to Body rolls, movement reduced) or wolf (striker form: +2 dice to Agility and Stealth, +1 band movement). You may speak but cannot use tools or cast other Rites while transformed.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Carrion Feast (9 XP)}

\textbf{Tags:} [CONSUME], [HEAL], [BOOST]

\textbf{Threshold:} The flesh of a fallen enemy, consumed raw.

Devour the strength of the fallen. Heal 1 Fatigue immediately, and gain +1 die to your next physical action. If the enemy was supernatural (spirit, demon, fey), you also gain a minor property of theirs for one scene.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Cycle's Crown (14 XP)}

\textbf{Tags:} [TRANSFORM], [NATURE], [BOOST], [PROTECT]

\textbf{Threshold:} A circlet of woven branches, antlers, and seasonal tokens.

Become an avatar of the cycle. Choose two: seasonal mastery (+2 dice to actions keyed to the current season), verdant command (command plant life in Near automatically), cycle's healing (remove one Condition or downgrade Harm by one level), or growth's protection (allies gain +1 Armor from natural cover). Begin a Seasonal Avatar [6] clock.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Seasonal Sense (4 XP)}

\textbf{Requirement:} Grimmir

You can predict the weather for the next hour with perfect accuracy. You know when the season will turn within a day. This is a passive ability.

\subsubsection{Wild Shape (6 XP)}

\textbf{Requirement:} Grimmir, Body 3+

Once per session, you may assume the form of a small beast (wolf, hawk, badger) for up to one scene. While transformed, you gain +2 dice to Stealth and Perception in natural terrain, but cannot speak or use tools. Mark 1 Fatigue after using this talent.

\subsubsection{Beast Speech (8 XP)}

\textbf{Requirement:} Grimmir, Spirit 3+

You can speak with animals. They are not compelled to obey, but they will answer simple questions truthfully—as they understand truth. The GM may have animals respond with limited knowledge or animal logic.

\subsection{Roleplaying a Root-Knight}

\begin{itemize}
\item You do not command nature. You listen to it. The forest speaks; you are not always fluent.
\item You may be called to settle disputes between villages and the wild—a wolf pack, a blighted grove, a tree that has grown across a road.
\item You may struggle with the cost of the hunt. Every kill is a sacrament. Every death feeds the earth. You do not kill lightly.
\item You may be asked to let something die—a forest, a creature, a memory. The cycle demands it.
\end{itemize}

\subsection{Sample NPC: Warden Thorne of the Root-Staff}

\textbf{Archetype:} Root-Knight

\textbf{Patron:} Grimmir

\textbf{Codex:} A living staff that grows leaves in spring

\textbf{Thiasos:} A wolf-ancestor who remembers the Runekeeper's grandfather

\textbf{Personality:} Quiet, patient, watchful. He has not spoken a word in seven years. The wolf speaks for him.

\textbf{Conflict:} A village wants to cut down an ancient oak that shelters the forest's heart. The oak is dying anyway. The cycle says: let it fall. The village says: save it. The wolf is silent.

\subsection{GM Guidance: Intrusions for Grimmir}

\begin{itemize}
\item A tree the Root-Knight passed yesterday now blocks their path—with a face carved into it that looks like theirs.
\item All food the Root-Knight touches spoils within an hour unless they offer the first bite to the earth.
\item A pack of wolves shadows the party for a full day. They do not attack. They simply watch.
\item The season turns early. The Root-Knight's seasonal boon shifts unexpectedly. The cycle is restless.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
