\section{Rites of the Wild Patrons}

The Wild patrons—Grimmir, the Carrion-King, and the Savage Heart—offer Rites that draw on the raw power of nature, decay, and predation. These Rites are physical, visceral, and often require the Runekeeper to shed blood, consume flesh, or stand bare-skinned before the elements.

\subsection{Grimmir — Root-Knight Rites}

\subsubsection{Low Rite: Rite of the Blood Offering (4 XP)}

\textbf{Tags:} [SACRIFICE], [NATURE], [STEALTH]

\textbf{Threshold:} A drop of your own blood (1 Fatigue) or the blood of a freshly killed animal.

The forest accepts your offering. For the remainder of the scene, you gain +1 die to all Survival and Stealth rolls in natural terrain. Additionally, you may ask the GM one yes/no question about a threat hidden in the wilderness (e.g., "Is there a predator watching us?").

\textbf{Push It:} The offering is accepted with hunger—you gain +2 dice instead, but you must leave a visible token (a hair, a scrap of clothing) on the ground. That token can be used by the GM later as a sympathetic link (spend 1 SB to have an animal or spirit track you).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Predator's Gaze (4 XP)}

\textbf{Tags:} [MARK], [FEAR], [TRACK]

\textbf{Threshold:} A tooth from a predator that killed its own prey.

Mark a single target as your prey. Gain +1 die on all rolls to track, pursue, or intimidate that target. The effect lasts until the target escapes or is killed.

\textbf{Push It:} The prey feels your gaze; they suffer -1 die to resist social pressure from you. Mark 1 SB (Spades).

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shifting Shape (8 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST]

\textbf{Threshold:} A pelt or mask from the beast whose form you take.

Assume the form of a predator for the scene. Choose bear (tank form: +1 Armor, +2 dice to Body rolls, movement reduced by one band) or wolf (striker form: +2 dice to Agility and Stealth, +1 band to movement, can Dash as a free action once per scene). Cultural variations: lion, leopard, tiger, or great cat where appropriate.

You may speak but cannot use tools or cast other Rites while transformed.

\textbf{Push It:} Shift as a free action once this scene, but mark 1 SB (Clubs) as the beast presses against your mind.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Carrion Feast (9 XP)}

\textbf{Tags:} [CONSUME], [HEAL], [BOOST]

\textbf{Threshold:} The flesh of a fallen enemy, consumed raw.

Devour the strength of the fallen. Heal 1 Fatigue immediately, and gain +1 die to your next physical action. If the enemy was supernatural (spirit, demon, fey, lycanthrope), you also gain a minor property of theirs for one scene.

\textbf{Push It:} Heal 1 Harm instead of Fatigue, but mark 1 SB (Hearts) as the feast attracts scavengers.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Cycle's Crown (14 XP)}

\textbf{Tags:} [TRANSFORM], [NATURE], [BOOST], [PROTECT]

\textbf{Threshold:} A circlet of woven branches, antlers, and seasonal tokens.

Become an avatar of the cycle. Choose two:
\begin{itemize}
\item \textit{Seasonal Mastery:} +2 dice to all actions keyed to the current season (spring: healing and growth; summer: strength and harvest; autumn: preservation and preparation; winter: endurance and hiding).
\item \textit{Verdant Command:} Command plant life in Near automatically (once per scene).
\item \textit{Cycle's Healing:} Remove one Condition or downgrade Harm by one level.
\item \textit{Growth's Protection:} Allies gain +1 Armor from natural cover.
\end{itemize}
Begin a Seasonal Avatar [6] clock.

\textbf{Push It:} Extend the blessings to all present; mark +2 Obligation and gain Seasonal Attunement (as per Grimmir's Corruption table).

\textbf{Obligation:} 3

\subsection{The Carrion-King — Grave-Warden Rites}

\subsubsection{Low Rite: Rite of the Rotting Harvest (5 XP)}

\textbf{Tags:} [DECAY], [HARVEST], [BOOST]

\textbf{Threshold:} Organic matter in early stages of decay, or the remains of a recently ended thing (burnt letter, wilted flower, shattered glass).

Accelerate natural decay to weaken or destroy an organic object (ropes, leather, wood), gaining +2 Effect on Break/Sabotage rolls against it. Alternatively, extract value from an ending: from a defeated enemy, gain +1 die on your next action; from a failed plan, reroll one 1 on your next roll; from a broken item, gain 1 Boon.

\textbf{Push It:} Spread decay to similar materials in Close range, or harvest additional value from an ending; in either case, mark 1 SB (Clubs) as the rot becomes noticeable or mark Fatigue 1 from dwelling on endings.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Borrowed Form (7 XP)}

\textbf{Tags:} [KNOWLEDGE], [SPIRIT], [BOOST]

\textbf{Threshold:} A token from a deceased being (hair, nail, written name).

Channel the essence of what was. Gain one skill die reflecting the deceased's expertise for one scene, or ask the GM one question about their knowledge or abilities.

\textbf{Push It:} The spirit's influence lingers—gain a permanent +1 die specialty (e.g., "Knowledge of the Dead Mason"), but suffer occasional possession-like effects at the GM's discretion.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Fertile Death (8 XP)}

\textbf{Tags:} [TRANSFORM], [AREA], [HEAL], [TERRAIN]

\textbf{Threshold:} Ashes, compost, or the remains of anything that once lived.

Transform death into growth: choose one—create beneficial terrain (cover, concealment, or advantageous positioning) OR grant allies +1 die to healing or recovery rolls for the scene.

\textbf{Push It:} Both effects apply, but the transformation attracts unwanted attention (vermin, scavengers, or curious onlookers); mark 1 SB (Spades).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Great Consumption (13 XP)}

\textbf{Tags:} [DECAY], [AREA], [SUMMON], [TRANSFORM]

\textbf{Threshold:} A significant amount of organic matter (corpse, fallen tree, collapsed building).

Transform a large area through decay and renewal. Choose two of the following:
\begin{itemize}
\item Create difficult terrain that favours you and your allies.
\item Summon a Cap 3 swarm of scavengers as temporary allies for the scene.
\item Generate valuable reagents worth 2 XP (usable for crafting or barter).
\end{itemize}

\textbf{Push It:} All three effects occur, but start a 6-segment Ecosystem Disruption clock that will cause problems later.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Eternal Cycle (14 XP)}

\textbf{Tags:} [TRANSFORM], [FATE], [SACRIFICE]

\textbf{Threshold:} The complete remains of something significant that has ended (a broken oath, a fallen kingdom's charter, a slain legendary beast).

Complete a transformation cycle. Destroy one major asset, enemy, or obstacle, and create something new of equal or greater value. The player and GM collaborate to define the transformation (e.g., a burned forest becomes fertile soil; a broken alliance becomes a new pact; a defeated warlord's corpse becomes a guardian tree).

\textbf{Push It:} The transformation is immediate and spectacular, but creates a 6-segment Cycle Debt clock. The King will demand another significant ending soon.

\textbf{Obligation:} 3

\subsection{The Savage Heart — Beast Warrior Rites}

\subsubsection{Low Rite: Rite of the Predator's Gaze (4 XP)}

\textbf{Tags:} [MARK], [FEAR], [TRACK]

\textbf{Threshold:} A tooth from a predator that killed its own prey.

Mark a single target as your prey. Gain +1 die on all rolls to track, pursue, or intimidate that target. The effect lasts until the target escapes or is killed.

\textbf{Push It:} The prey feels your gaze; they suffer -1 die to resist social pressure from you. Mark 1 SB (Spades).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Blood Scent (5 XP)}

\textbf{Tags:} [SENSE], [FEAR], [TRACK]

\textbf{Threshold:} A fresh kill's blood smeared on your face.

You scent the air like a beast. Gain +2 dice to Investigation or Survival when tracking by smell. You may also detect fear—you sense if a creature within Near is Frightened.

\textbf{Push It:} Your sense expands to Far, but you are overwhelmed by the smell of blood; mark 1 SB (Hearts) and suffer -1 die to social rolls until the end of the scene.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shifting Shape (8 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST]

\textbf{Threshold:} A pelt or mask from the beast whose form you take.

Assume the form of a predator for the scene. Choose bear (tank form: +1 Armor, +2 dice to Body rolls, Guardian's Pounce: once per scene when you protect an ally, you may also knock the attacker prone) or wolf (striker form: +2 dice to Agility and Stealth, +1 band movement, Ambush Predator: when attacking from stealth, gain +1 Effect and ignore 1 point of Armor). Cultural variations: lion, leopard, tiger, great cat as appropriate.

You may speak but cannot use tools or cast other Rites while transformed.

\textbf{Push It:} Shift as a free action once this scene, but mark 1 SB (Clubs) as the beast presses against your mind.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Pack Hunt (7 XP)}

\textbf{Tags:} [TEAMWORK], [BEAST], [BOOST]

\textbf{Threshold:} A token from each member of your pack (a claw, a feather, a scrap of hide).

You and up to three allies form a hunting pack. While within Near of each other, all gain +1 die to attacks against a single designated prey, and the first ally to attack each round gains +1 Effect.

\textbf{Push It:} Coordination extends to Far, but hunters become territorial; mark 1 SB (Spades) as rivalries flare.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Savage Transformation (13 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST], [FEAR]

\textbf{Threshold:} The complete pelt of an apex predator, worn during the rite.

Fully transform into an apex predator for one scene. You gain the physical attributes of the beast: Body +2, +2 Armor from thick hide, and your melee attacks deal +2 Harm. You cannot speak or use tools. You may end the transformation early as an action. When the transformation ends, you mark 2 Fatigue.

\textbf{Push It:} The transformation becomes contagious—nearby allies may also transform but risk losing their humanity; mark +2 Obligation.

\textbf{Obligation:} 3

\subsection{Rite Synergies}

Some Rites work well together for Wild patrons:

\begin{itemize}
\item \textbf{Rite of the Blood Offering + Rite of the Predator's Gaze:} Offer blood to mark your prey. The forest accepts the offering; the hunt is sanctified.
\item \textbf{Rite of the Carrion Feast + Rite of the Great Consumption:} Devour the fallen, then consume the battlefield. Death feeds death. The cycle continues.
\item \textbf{Rite of the Shifting Shape (Savage Heart) + Rite of the Pack Hunt:} Transform into a predator and lead your pack. The hunt is no longer yours alone—it belongs to the pack.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
