\section{The Savage Heart — Predation and the Red Maw}

\subsection{Domain and Epithets}

The Savage Heart is the truth behind the green veil: nature not as garden, but as gullet. It is the claw that tears, the fang that crushes, the hunger that stalks without malice and kills without mercy. Its followers are not gentle guardians of the wild—they are its teeth and claws. Power comes from taking, from proving strength, from the raw act of predation.

The Savage Heart does not ask. It takes. The deer does not weep for the wolf. The sapling does not beg the axe. In the red maw, there is no cruelty—only truth. The truth that all flesh feeds the earth, and the earth feeds the strong.

\subsection{Cultural Origins}

The Savage Heart wears the face of the apex predator in every land:

\begin{itemize}
\item In the \textbf{Amaranthine core}, it is the Great Bear, patient and unstoppable, a wall of fur and fury.
\item Among the \textbf{Ykrul} and \textbf{Kuvani}, it is the Wolf of the Steppe, a pack hunter with tireless stamina.
\item In \textbf{Oshiira}, it is the Golden Lion, whose roar shatters the hunt's silence.
\item In \textbf{Sekogo}, it is the Leopard of the Tall Grass, a shadow that strikes once and is gone.
\item In \textbf{Dhahara}, it is the Tiger of the Mangroves, whose amber gaze promises a death that runs faster than you can flee.
\end{itemize}

\subsection{Warrior Archetype: The Beast Warrior}

A Runic Warrior of the Savage Heart is not a gardener of rot or a warden of seasons. They are a \textit{Beast Warrior}—a predator who wears the skin of their kills, who hunts as sacrament, who transforms into the very thing they stalk. Their Codex is a predator's tooth or a pelt worn as a cloak. Their Thiasos is a wolf-ancestor, a great cat's spirit, or a sentient blade carved from the fang of a legendary beast.

\subsection{Codex Form}

The Beast Warrior's Codex is a trophy of the hunt:

\begin{itemize}
\item \textbf{Predator's Tooth:} A large canine tooth on a leather thong. It is warm when prey is near. It grows heavier with each kill.
\item \textbf{Pelt Cloak:} The cured hide of an apex predator (bear, wolf, lion, tiger). Wearing it grants +1 die to Intimidation.
\item \textbf{Claw Necklace:} A string of claws from worthy kills. Each claw holds a memory of the hunt. The necklace clicks when the hunt begins.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Wolf-Ancestor:} A grey wolf the size of a pony, with eyes that reflect the Runekeeper's first kill. It tracks across any terrain and fights beside its pack-brother.
\item \textbf{Great Cat Spirit:} A translucent leopard or tiger that moves without sound. It can stalk a target for hours without being seen.
\item \textbf{The Fang (Sentient):} A blade carved from the tooth of a legendary beast. It speaks in growls. It hungers. It demands to be fed.
\end{itemize}

\subsection{Core Mechanic: The Hunt}

When the Beast Warrior marks a target as prey (an action, requires line of sight or a successful tracking roll), they gain +1 die to all rolls to track, pursue, or confront that target. The mark lasts until the target escapes or is killed.

When the Beast Warrior kills their marked prey, they may immediately clear 1 Fatigue and gain 1 Boon. This Boon may only be used on physical actions.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Predator's Gaze (4 XP)}

\textbf{Tags:} [MARK], [FEAR], [TRACK]

\textbf{Threshold:} A tooth from a predator that killed its own prey.

Mark a single target as your prey. Gain +1 die on all rolls to track, pursue, or intimidate that target. The effect lasts until the target escapes or is killed.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Blood Scent (5 XP)}

\textbf{Tags:} [SENSE], [FEAR], [TRACK]

\textbf{Threshold:} A fresh kill's blood smeared on your face.

You scent the air like a beast. Gain +2 dice to Investigation or Survival when tracking by smell. You may also detect fear—you sense if a creature within Near is Frightened.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shifting Shape (8 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST]

\textbf{Threshold:} A pelt or mask from the beast whose form you take.

Assume the form of a predator for the scene. Choose bear (tank form: +1 Armor, +2 dice to Body rolls, movement reduced) or wolf (striker form: +2 dice to Agility and Stealth, +1 band movement). You may speak but cannot use tools or cast other Rites while transformed.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Pack Hunt (7 XP)}

\textbf{Tags:} [TEAMWORK], [BEAST], [BOOST]

\textbf{Threshold:} A token from each member of your pack (a claw, a feather, a scrap of hide).

You and up to three allies form a hunting pack. While within Near of each other, all gain +1 die to attacks against a single designated prey, and the first ally to attack each round gains +1 Effect.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Savage Transformation (13 XP)}

\textbf{Tags:} [TRANSFORM], [BEAST], [BOOST], [FEAR]

\textbf{Threshold:} The complete pelt of an apex predator, worn during the rite.

Fully transform into an apex predator for one scene. You gain the physical attributes of the beast: Body +2, +2 Armor from thick hide, and your melee attacks deal +2 Harm. You cannot speak or use tools. You may end the transformation early as an action. When the transformation ends, you mark 2 Fatigue.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Beast Form (6 XP)}

\textbf{Requirement:} The Savage Heart, Body 3+

Once per session, you may partially transform into a predator for one scene. Choose one: gain +2 dice to Stealth (leopard/tiger), +2 dice to Intimidation (bear/lion), or +1 band to movement (wolf/hound). You cannot use this talent while already transformed.

\subsubsection{Blood Hunt (4 XP)}

\textbf{Requirement:} The Savage Heart

When you are tracking a target that you have wounded, you gain +1 die to all tracking and pursuit rolls. The scent of their blood is a trail you cannot lose.

\subsubsection{Alpha's Roar (8 XP)}

\textbf{Requirement:} The Savage Heart, Presence 3+

Once per session, you may make a terrifying sound—a roar, a howl, a scream. All enemies within Near range must test Resolve (DV 4) or become Shaken (-1 die to all actions) for one exchange. This does not affect creatures that are immune to fear.

\subsection{Roleplaying a Beast Warrior}

\begin{itemize}
\item You do not fight for glory. You hunt for survival. The kill is sacrament; the meat is blessing.
\item You may be called to track fugitives, hunt dangerous beasts, or lead a pack in war.
\item You may struggle with the hunger. The Savage Heart does not ask you to be gentle. It asks you to be true—to the hunt, to the pack, to yourself.
\item You may be asked to kill something that does not deserve to die. The Heart does not care about deserving. The Heart is hungry.
\end{itemize}

\subsection{Sample NPC: Mkara of the Leopard's Shadow}

\textbf{Archetype:} Beast Warrior

\textbf{Patron:} The Savage Heart

\textbf{Codex:} A leopard's tooth on a leather thong

\textbf{Thiasos:} A great cat spirit that moves without sound

\textbf{Personality:} Silent, patient, deadly. She has never lost a hunt. She has never needed to speak.

\textbf{Conflict:} A child has wandered into the territory of a wounded leopard. The leopard is dangerous. The child is innocent. The hunt says: kill. The heart says: wait. The leopard is watching.

\subsection{GM Guidance: Intrusions for The Savage Heart}

\begin{itemize}
\item The Beast Warrior's prey escapes—not because of poor skill, but because the prey was not worthy. The Savage Heart is displeased.
\item A rival predator enters the Beast Warrior's territory. It does not attack. It watches. It is judging.
\item The Beast Warrior's pelt cloak begins to itch. The beast whose skin it was is not entirely dead. It dreams.
\item The hunt lasts too long. The Beast Warrior marks 1 Fatigue. The hunger grows. The Savage Heart whispers: \textit{faster}.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
