\part{The Southern Realms (Akilan)}
% After \part{The Southern Realms (Akilan)}
\begin{quote}
  \small
  \textit{``The ledger does not forgive arithmetic errors. A single miscounted bushel can unmake a house, a tribe, an empire. I am not here to be kind. I am here to be correct. But even the Widow's ink runs when the monsoon weeps.''}
  \noindent --- Dusana of the Raven Road, at the weigh-house of Mbaro
  \end{quote}

% ------------------------------
% CHAPTER: OSHIIRA
% ------------------------------
\chapter{Oshiira}
\emph{The Ledger Empire.}

\noindent
{\Large
\textit{The ledger does not forgive arithmetic errors. A single miscounted bushel can unmake a house, a tribe, an empire. I am not here to be kind. I am here to be correct.}

\vspace{0.5em}

I am a Grain Inspector. My stamp leaves a mark that glows under moonlight. I know.

\vspace{0.5em}

Oshiira is a confederacy of hundreds of tribes and city-states, united by the Grain Ledger. The capital, Mbaro, is a city of canals, weigh-houses, and the great Granary Tower. Oshiira is the breadbasket of the Amaranthine; its millet and sorghum fill the holds of Ecktorian, Vhasian, and Silkstrand ships. The empire operates through a bureaucracy as complex as any in the north---weighing-houses, canal-mothers, and ledgers that track every grain shipment with mathematical precision. But here's the secret the Widow doesn't want you to know: the Ledger has a page that has been torn out. That page named the true heir. No one knows where it is. Everyone is looking.

\vspace{0.5em}

The Vermilion Corps audits not only grain but also loyalty. The Narrow Captains provoke Kahfagia just to remind the Admiralty who really controls the straits. And the Widow---the empress who has held the confederacy together for centuries---is dying. Paranoia is slowly replacing wisdom. Her four children circle the throne like vultures. The Vermilion Corps maintains order, but no one knows which heir the Corps truly serves. Not even the Corps itself.

\vspace{0.5em}

Never speak of succession before naming the Widow. The empress's health is a state secret, and the Vermilion Corps has ears in every weigh-house. Speak carelessly, and your name may be the next one erased from the Ledger.

\vspace{1em}

\noindent\textit{--- Vermilion Corps Auditor, who counts her own breaths to ensure she is not lying}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Millet porridge with dried fish---the millet is grown in the flood plains of the Sahel, the fish is smoked over canal-barges. The porridge is always the same. The fish varies by season. The seasons are measured by the height of the river. The river is the Widow's blood. The Widow's blood is low.

\textbf{What does a child learn before they can walk?} To weigh a grain. Oshiira children are given a balance scale and a single grain of millet. By age five, they can tell you the weight difference between a grain harvested in the northern Sahel and a grain harvested in the southern riverlands. The difference is the tax. The tax feeds the empire. The empire is hungry.

\textbf{What is the first thing you notice about a stranger?} Their hands. Oshiira merchants have ink-stained fingers---the ink is from the Ledger. A stranger whose fingers are clean has never signed a contract. A stranger whose fingers are stained black has signed too many. A stranger whose fingers are stained red has signed a death warrant. The red ink is not ink. The red ink is blood.

\textbf{What is the most common crime?} Ledger fraud---but ledger fraud is not a crime, it is a career. The Vermilion Corps audits the auditors. The auditors audit the clerks. The clerks audit the merchants. The merchants audit the farmers. The farmers audit the land. The land does not lie. The land does not need to.

\textbf{What does no one talk about but everyone knows?} The Widow has not been seen in public for seven years. Her voice is a recording. Her seal is a stamp. Her decrees are written by the Vermilion Corps. The Corps signs in her name. The Corps speaks in her voice. The Corps is the Widow. The Widow is dead. The empire continues.

\textbf{Where do people go when they die?} To the Granary Tower---the great stone silo at the center of Mbaro. The dead are not buried. They are ground into meal and mixed with the millet. The grain is shipped across the Amaranthine. The dead feed the empire. The empire does not know. The empire does not ask.

\textbf{How do you apologize for a serious wrong?} You submit a correction to the Ledger. The correction must be witnessed by a Vermilion Corps auditor. The auditor stamps the correction with their seal. The seal glows under moonlight. The moonlight is the Widow's eye. The Widow sees everything. The Widow forgives nothing.

\textbf{What is the most important festival?} The Harvest of the Ledger (autumn). The Ledger is opened in the great square. The names of the year's debtors are read aloud. The debtors are given until dawn to pay. The debtors who cannot pay are added to the Ledger's ninth page. The ninth page is sealed with wax. The wax is black. The black is mourning.

\textbf{What is the secret that could destroy a powerful person?} The Ledger's missing page is not missing. It was burned by the Widow herself, in the year the last heir died. The page's ash was mixed into the mortar of the Granary Tower. The ash still glows under moonlight. The glow spells the heir's name. The name is the Widow's own. The Widow killed her child. The empire does not know. The empire would not care.
\end{almanac}

\newpage
\input{../../generators/oshiira.tex}
\newpage

% ------------------------------
% CHAPTER: ASHAAN
% ------------------------------
\chapter{Ashaan}
\emph{The Veiled Throne.}

\noindent
{\Large
\textit{The Veil is not a choice. It is survival. You will show your eyes and nothing more. Speak in thirds---offer two options and a hidden third. The ninth word is always a knife.}

\vspace{0.5em}

I am a Breath-less Agent. They say Ikasha sleeps. I say she is patient. The Sisters have ruled for generations; they believe the Veil is eternal. But the Veil is silk. Silk frays. And I am learning to bite through---one whisper at a time.

\vspace{0.5em}

Ashaan is a nation that died long ago. The Veiled Sisters keep its corpse walking through spirit-binding, terror, and the endless consumption of their own people. The capital, Galanina, is a city of black marble and brass cages. The Red Desert, which now covers half the country, was once fertile land; the Sisters' magic leeched the life from it. There is no recovery, only a slower dying. But here's the secret the Sisters don't want you to know: the Veil is cracking. The spirits they've bound for centuries are whispering to each other. And the Hollow has begun to notice.

\vspace{0.5em}

The Esoiti control every ledger and every gate. The Black Hand hunts without uniform---but the Hand is human. They were once bakers, clerks, mothers, soldiers. Some are fanatics who chose the knife. Others are desperate souls who were given a choice: serve or vanish. They do not sleep well. Their hands shake when the Sisters'' acolytes pass. And the Sisters---the Sisters are not human. They do not negotiate. They do not threaten. They state what will be. Their voices are the same as the wind over a mass grave. You cannot bribe a divine disaster. You can only survive its attention.

The Breath-less meet in root cellars, learning to read ledgers without leaving marks. We are not heroes. We are the ones who learned that a whisper can be louder than a scream---and that the Sisters fear silence more than they fear rebellion. Silence means the Veil is healing. Silence means the bound are forgetting. The Sisters cannot abide forgetting.

\vspace{0.5em}

Never name the dead. The Sisters bind spirits to names. Speak a dead name aloud, and you may find yourself bound alongside them.

\vspace{1em}

\noindent\textit{--- Tamar, Breath-less cell leader, who has three names---each for a different jurisdiction}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Black bread and thin soup---the bread is made from grain grown on the edge of the Red Desert, the soup is flavoured with salt and memory. The memory is of green fields. The fields are gone. The bread is bitter.

\textbf{What does a child learn before they can walk?} To veil their face. Ashaan children are taught to cover their mouths when speaking, their eyes when listening, their ears when hearing secrets. A child who cannot veil is a child who will be taken by the Sisters. The Sisters take children. The children become spirits. The spirits serve the Sisters. The Sisters never stop taking.

\textbf{What is the first thing you notice about a stranger?} Their veil. Ashaan veils are coded---red for mourning, white for marriage, black for the Sisters. A stranger whose veil is gray is a Breath-less agent. A stranger whose veil is torn is a fugitive. A stranger with no veil is a fool. Fools are collected by the Black Hand. The Black Hand does not ask questions. The Black Hand was once a person. The Black Hand no longer remembers who.

\textbf{What is the most common crime?} Spirit-trading---but spirit-trading is not a crime, it is the economy. The Sisters bind spirits to brass cages. The cages are sold to the Esoiti. The Esoiti sell the spirits to foreign buyers. The foreign buyers do not know what they are buying. The Sisters do not care. The Sisters are not merchants. They are a natural disaster with a ledger.

\textbf{What does no one talk about but everyone knows?} \textit{Everyone is afraid.} The Black Hand fears the Esoiti's audits. The Esoiti fears the Sisters`` gaze. The Sisters fear the Hollow's patience. The baker fears the guard. The guard fears the Hand. The Hand fears the name on tomorrow's list. The only person in Ashaan who is not afraid is the child who has not yet learned to read the ledger. That child will be afraid by next spring.

\textbf{Where do people go when they die?} To the brass cages. The dead are not buried. They are bound into spirit-lanterns and hung from the walls of Galanina. The lanterns glow green. The green is the light of the bound. The bound cannot speak. The bound cannot scream. The bound can only watch. Sometimes they weep. The tears evaporate into glass.

\textbf{How do you apologize for a serious wrong?} You offer a name---a dead name, a forgotten name, a name the Sisters do not yet know. The Sisters bind the name to a brass cage. The cage is placed in the Hall of Silence. The Hall is full. The Hall is always full. The Hall is never silent. The cages whisper. The whispers are the names of those who were erased.

\textbf{What is the most important festival?} The Night of the Unbound (winter solstice). The spirits strain against their cages. The Sisters renew their bindings. The ceremony takes three days. The city is silent for three days. The silence is not prayer. The silence is fear. If you listen closely during the silence, you can hear the Hollow breathing.

\textbf{What is the secret that could destroy a powerful person?} The High Sister is not a sister. She is a spirit---the first one bound, nine hundred years ago. She bound herself to the throne. She cannot leave. She cannot die. She cannot stop binding. The other Sisters do not know. The other Sisters are her prisoners. The prisoners guard the jailer. The jailer does not sleep. Her veil is made of gold coins from a dozen dead empires. The coins do not clink. They weep when she speaks.
\end{almanac}

\newpage
\input{../../generators/ashaan.tex}
\newpage

% ------------------------------
% CHAPTER: NGOMEBE
% ------------------------------
\chapter{Ngomebe}
\emph{The Stone Men.}

\noindent
{\Large
\textit{A city that stops moving is a city that has chosen to die. The stone remembers every mile, every roller, every oath sworn in the shadow of the keystone. We do not build cities. We raise them from the earth and teach them to walk.}

\vspace{0.5em}

I am a Stone-Cutter. My grandfather carved the keystone for this city. My father greased the rollers. I inherited a sledge and a song. The song is the city's heartbeat. If I stop singing, the city stops walking---and so do I.

\vspace{0.5em}

Ngomebe are a people of deep, iron-dark skin and older ties to the southern ranges. Their cities walk---colossal stone structures on rollers and sledges, dragged by teams of oxen, magic, and Aeler-designed capstans. A walking city is not a curiosity; it is a living argument against stillness. But here's the secret the Mason-Kin don't want you to know: the stone is not just a material. It listens. It remembers. And sometimes, it disagrees. When the keystone cracks, it is not a flaw---it is a message. The ancestors are trying to tell you something. The question is whether you're brave enough to read the cracks.

\vspace{0.5em}

The Root-binders plant acorns in the roller-pits, singing to the earth. They believe the cities should stop and let the forest reclaim the stone. The Aeler engineers maintain the bearings, but they are neither masters nor servants---they are partners in a pact older than memory. And the Ancestral Healers speak for the dead, who vote through the cracks in the keystone. Their votes are always unanimous. Their verdicts are always final.

\vspace{0.5em}

Never carve a name into living rock. The stone remembers. It will whisper that name to every spirit of the earth, and the debt will follow your bloodline for seven generations.

\vspace{1em}

\noindent\textit{--- Remi, Stone-Cutter, whose city left him behind}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Stone-ground meal and fermented milk---the meal is ground between the city's own millstones, the milk is from goats that walk alongside the city. The goats are older than the city. The goats remember where the city has been. The goats do not speak. The goats do not need to.

\textbf{What does a child learn before they can walk?} To sing the city's song. Ngomebe children are taught the rhythm of the rollers---eight beats to a mile, eight miles to a day, eight days to a season. The ninth beat is the city's secret. The ninth beat is never sung aloud.

\textbf{What is the first thing you notice about a stranger?} Their dust. Ngomebe are coated in fine stone-dust---from the quarries, the rollers, the wind that follows the city. A stranger whose dust is white has come from the mountains. A stranger whose dust is red has come from the desert. A stranger with no dust has never walked with a city. The cities do not trust the dustless.

\textbf{What is the most common crime?} Stone-theft---but stone-theft is not a crime, it is a wound. The stone is the city's memory. A stolen stone is a forgotten year. The year is forgotten by everyone who walked it. The forgetfulness spreads. The city's song stutters. The stutter is a warning.

\textbf{What does no one talk about but everyone knows?} The city is not moving toward anything. It is moving away. The Ngomebe are fleeing something that lies to the south. The thing in the south is older than the stone. The thing in the south does not walk. It waits.

\textbf{Where do people go when they die?} Into the keystone. The dead are not buried. Their ashes are mixed into mortar and pressed into the city's foundation. The dead become part of the city's weight. The city walks heavier each year. The dead do not complain. The dead are the city's anchor.

\textbf{How do you apologize for a serious wrong?} You carve a new roller. The roller must be cut from a single stone, with no cracks, no flaws, no forgotten names. The roller is placed beneath the city's foundation. If it rolls true, you are forgiven. If it cracks, you are not. If it crumbles, you are added to the mortar.

\textbf{What is the most important festival?} The Turning of the Rollers (spring equinox). The city stops for one day. The rollers are inspected, greased, and blessed. The city's song is sung from dawn to dusk. The song has eight verses. The ninth verse is silence. The silence is for the dead. The dead are listening.

\textbf{What is the secret that could destroy a powerful person?} The city is not a city. It is a tomb---a monument to a king who died before the first stone was laid. The king's name is carved into the keystone. The keystone is at the city's heart. The heart is inaccessible. The king is patient. The king is waiting to be remembered.
\end{almanac}

\newpage
\input{../../generators/ngomebe.tex}
\newpage

% ------------------------------
% CHAPTER: TAHARKA
% ------------------------------
\chapter{Taharka}
\emph{The Monsoon Crown.}

\noindent
{\Large
\textit{The passes remember every army that tried to climb them. The mountains do not forgive haste, and they do not forget a foreign banner. We do not guard the highlands. We listen to them. They tell us who to let through.}

\vspace{0.5em}

I am a Horse Trader. Before you praise the rider, praise the horse. Before you praise the horse, praise the mountain that grew its lungs. A Taharkan foal is born breathing thin air. That is why our cavalry has never been broken.

\vspace{0.5em}

Taharka grows rich on monsoon trade---spices, coffee, and incense that travel east to the distant ocean and west to the Gulf. The kingdom has never been conquered; its passes are a graveyard for Ashaani armies and a barrier that Oshiiran grain barges cannot cross. The highlands are a fortress of rock and thin air, where every valley speaks a different dialect and every ridge has a watchtower. But here's the secret the Crown Council doesn't want you to know: the passes are alive. They have a memory. They remember every army that tried to climb them, every foreign banner that flew too high, every oath that was broken at the summit. And they do not forgive.

\vspace{0.5em}

The Ridge-Watchers are the eyes of the kingdom, stationed in rock-chapels that double as signal towers. The horse is not a mount; it is a lineage, a treaty, a weapon. To control the highland breeds is to control the passes---and to control the passes is to remain free. The lowland coast operates under a different law: the ujamaa of the dhows, a mutual aid network where every captain owes a share of every voyage to the community. A captain who cheats the ujamaa finds his ship's name struck from every mooring post in the archipelago.

\vspace{0.5em}

Never refuse a second cup of tea. The first is hospitality; the second is a test of honor. The third is a promise. Refuse the second, and you have insulted the house---and the house's ancestors, who are listening through the floorboards.

\vspace{1em}

\noindent\textit{--- Ridge-Watcher Captain Anwar, who has not descended below the snow line in twenty years}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Barley flatbread and spiced tea---the bread is baked on iron griddles, the tea is brewed with cardamom and mountain herbs. The herbs are harvested from passes that only the Ridge-Watchers know. The Ridge-Watchers do not share.

\textbf{What does a child learn before they can walk?} To read the wind. Taharkan children are taught the eight winds by name: North, South, East, West, Monsoon, Dry, Storm, and the ninth---the wind that has no name. The ninth wind is the mountain's breath. The mountain's breath is death.

\textbf{What is the first thing you notice about a stranger?} Their breath. The air is thin in Taharka. A stranger who breathes too fast has not acclimated. A stranger who breathes too slow is hiding something. A stranger who does not breathe at all is a spirit. The spirits are not to be spoken to.

\textbf{What is the most common crime?} Horse theft---but horse theft is not a crime, it is a declaration of war. The horse is the kingdom's bloodline. A stolen horse is a stolen ancestor. The thief's clan is responsible. The clan pays in blood or in trade. The trade is never enough. The blood is always too much.

\textbf{What does no one talk about but everyone knows?} The Crown Council has not met in ten years. The passes have been closed by snow, by war, by the simple fact that the councillors cannot agree on a meeting place. The kingdom is ruled by the Ridge-Watchers. The Ridge-Watchers are not elected. The Ridge-Watchers are the mountains' voice. The mountains are silent.

\textbf{Where do people go when they die?} To the high passes. The dead are laid on stone platforms, wrapped in woollen shrouds. The wind takes them. The wind carries their names to the peaks. The peaks remember. The peaks are the kingdom's cemetery.

\textbf{How do you apologize for a serious wrong?} You offer a horse---the best horse in your stable. The horse is given to the wronged party. If they accept, the debt is paid. If they refuse, the debt is doubled. If they kill the horse, the debt is blood. Blood is never repaid.

\textbf{What is the most important festival?} The First Rain (monsoon's arrival). The passes open for the first time in months. The caravans descend from the highlands. The lowland ports receive the first spice ships. The celebration lasts nine days. The ninth day is silent. The silence is for the dead of the passes.

\textbf{What is the secret that could destroy a powerful person?} The Crown Council's missing member is not missing. He is buried under a landslide that was not an accident. The landslide was caused by the Ridge-Watchers. The councillor had learned the passes' secret: they are not natural. They were carved by something older. The something is still there. The Ridge-Watchers are its servants.
\end{almanac}

\newpage
\input{../../generators/taharka.tex}
\newpage

% ------------------------------
% CHAPTER: SEKOGO
% ------------------------------
\chapter{Sekogo}
\emph{The Green Council.}

\noindent
{\Large
\textit{The jungle has an immune system. When a fever rises, we cut it out. When an oath is broken, we collect the price. We are not cruel. We are necessary. The tree remembers every unpaid debt, and we are its teeth.}

\vspace{0.5em}

I am a Honey-Gatherer. The bees know my name. They do not sting me because I sing to them the song of the first blossom. My grandmother taught me that the honey is not a gift; it is a loan. One day, the grove will ask for it back.

\vspace{0.5em}

Sekogo lies between the Sahel grasslands and the southern plateau---a dry savannah that grows wetter as you travel east, until the baobabs give way to deep jungle and the air turns thick with the smell of damp earth and overripe fruit. Here, the Green Council rules through consensus, honey-orchards are sacred, and the old spirits still walk. Slavery is forbidden; the jungle itself would rise against any slaver who crossed its border. But here's the secret the Council doesn't want you to know: the jungle has a favorite. It is not the Conservators who protect the old pacts. It is not the Wakers who want to rouse the jungle with fever-mists and bloodier rites. It is the Ukwe---the Speaking Tree---that decides whose debts are collected and whose are forgiven. And the Ukwe does not answer to the Council.

\vspace{0.5em}

The Leopard Society hunts those who break the deepest taboos. The Earth Society keeps the Calendar of Debt, a sacred list written on bark and buried at the Ukwe's roots. The nganga work with minkisi---power objects inhabited by spirits---hammering nails into carved figures to release magic. Each nail is a new debt to the spirit within. Trickster spirits---children of Ananse, the spider-trickster---weave illusions and swap debts for laughter. And in the oldest groves, the Ninki Nanka---a serpent with a mirror-back---slithers between roots, its reflection showing the lies you have told.

\vspace{0.5em}

Never cut a marked tree. Each such tree is a living witness. The bark grows over the names of debtors, but the wood remembers. Cut it, and the Ukwe will add your name to its ledger.

\vspace{1em}

\noindent\textit{--- Hunter Okoro, Leopard Society mask-wearer, who never removes his mask}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Mashed yams and groundnut stew---the yams are grown in clearings that shift with the seasons, the groundnuts are harvested by monkey-tribes who speak the old tongue. The monkeys are not paid. The monkeys are remembered. Memory is currency.

\textbf{What does a child learn before they can walk?} To name the spirits. Sekogo children are taught the eight spirits of the grove: River, Tree, Stone, Rain, Fire, Wind, Honey, and the ninth---the spirit that has no name. The ninth spirit is the Ukwe's shadow. The shadow watches. The shadow never forgives.

\textbf{What is the first thing you notice about a stranger?} Their mask. Every Sekogo wears a mask---carved wood, painted with the symbols of their lineage. A stranger whose mask is cracked has broken a taboo. A stranger whose mask is blank has no lineage. A stranger with no mask is not a person. The Leopard Society hunts the maskless.

\textbf{What is the most common crime?} Debt-forgetting---but debt-forgetting is not a crime, it is a sickness. The Ukwe keeps the Calendar of Debt. A name that is forgotten is a debt that is unpaid. The unpaid debt grows roots. The roots strangle the debtor's house. The house collapses. The jungle reclaims it.

\textbf{What does no one talk about but everyone knows?} The Green Council does not rule. The Ukwe rules. The Council is the Ukwe's voice---eight speakers chosen by the tree's roots. The ninth seat is empty. The ninth speaker was the tree itself. The tree does not speak. The tree waits.

\textbf{Where do people go when they die?} To the Ukwe's roots. The dead are not buried. Their bodies are laid at the base of the Speaking Tree. The roots absorb them. The tree grows a new leaf for each soul. The leaves whisper the names of the dead. The whisper is the wind.

\textbf{How do you apologize for a serious wrong?} You hammer a nail into a nkisi---a power figure carved by the nganga. The nail represents your wrong. The spirit inside the nkisi accepts your apology. The nail rusts. When the nail is gone, your debt is paid. The nail never rusts.

\textbf{What is the most important festival?} The Honey Harvest (autumn). The bees are sung to sleep. The hives are opened. The honey is collected in copper bowls. The first bowl is offered to the Ukwe. The tree drinks. The tree does not thank.

\textbf{What is the secret that could destroy a powerful person?} The Ukwe is not a tree. It is a prison---the binding of a spirit that tried to consume the jungle nine centuries ago. The spirit is still alive. The tree is its cage. The roots are its chains. The Green Council is its jailers. The jailers have forgotten what they guard.
\end{almanac}

\newpage
\input{../../generators/sekogo.tex}
\newpage

% ------------------------------
% CHAPTER: LETHAI-AR OF THE CRIMSON VALLEY
% ------------------------------
\chapter{Lethai-ar of the Crimson Valley}
\emph{Silk Bridges and Web-Law.}

\noindent
{\Large
\textit{The web-law is not written on parchment. It is woven into every bridge, every strand, every knot. When a promise is made, the valley remembers. When a promise is broken, the silk tightens. We are not judges. We are the loom.}

\vspace{0.5em}

I am a Mark-Cutter. Each line on my skin is a promise I witnessed. Each line is a debt I carry. When I die, they will cut these marks from my body and bury them at the Ukwe's roots. Then the tree will remember what I owed. I hope it is enough.

\vspace{0.5em}

The Crimson Valley is a deep jungle rift between Sekogo and Oshiira---a wound of red stone, carved by a river that runs rust-colored after every rain. Here, the Lethai-ar guard their borders with fierce neutrality, their chestnut-brown skin blending with the canyon walls. Silk bridges span the gorges, woven from spider-sap and oath-hair, strong enough to hold a caravan and light enough to sway in the monsoon wind. But here's the secret the wardens don't want you to know: the web-law is not unwritten because it is ancient. It is unwritten because it changes. Every warden carries a different clause in memory. Together, they form a complete web---but the web is never complete for long. The valley is learning. And it is not learning to be kind.

\vspace{0.5em}

The Lethai-ar are adopted children of the old powers---not fallen, but chosen to carry memory. Each can name the specific mountain, river, or grove where their ancestor left the spirit world to become mortal. Their skin-tattoos are not decoration; they are a public ledger, written on skin, read by anyone who knows the phase of the crimson moon. A Lethai-ar who bears a Mark they cannot fulfil loses standing. A warden who cuts a bridge without cause is exiled.

\vspace{0.5em}

Never step onto a silk bridge without first naming a witness. The bridge will remember your footsteps. If you fall, it will ask why no one was watching.

\vspace{1em}

\noindent\textit{--- Silk-Warden Anara, whose left arm is covered in Marks from wrist to shoulder}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge and fermented tea---the rice is grown on terraces cut into the canyon walls, the tea is aged in clay pots buried beneath the bridges. The tea is bitter. The bitterness is for memory.

\textbf{What does a child learn before they can walk?} To read a Mark. Lethai-ar children are taught the eight Mark-shapes: Promise, Debt, Witness, Apology, Interest, Cut, Forgive, and the ninth---the empty space where a Mark was erased. The erased Mark is a broken promise. The broken promise is a cut. The cut is a scar.

\textbf{What is the first thing you notice about a stranger?} Their Marks. A stranger whose Marks are fresh has made new promises. A stranger whose Marks are faded has unpaid debts. A stranger whose Marks are cut has been exiled. The exiled are not spoken to. The exiled are not remembered.

\textbf{What is the most common crime?} Bridge-cutting---but bridge-cutting is not a crime, it is a war. The bridge is a promise between two shores. Cutting it is breaking that promise. The valley responds. The valley tightens its web. The web is the law. The law is the valley.

\textbf{What does no one talk about but everyone knows?} The valley is not a place. It is a creature---a sleeping serpent whose spine is the canyon walls, whose scales are the silk bridges, whose heart is the rust-colored river. The Lethai-ar are its parasites. They live on its back. They weave its silk. They do not know it is asleep.

\textbf{Where do people go when they die?} To the river. The dead are not buried. Their bodies are wrapped in silk and cast into the rust-colored water. The river carries them downstream, into the Ukwe's territory. The Ukwe drinks. The Ukwe remembers. The Ukwe does not forget.

\textbf{How do you apologize for a serious wrong?} You offer a Mark---a promise cut from your own skin. The Mark is pressed into the wronged party's ledger. If they accept, the debt is transferred. If they refuse, the Mark is buried. Buried Marks are debts that never die.

\textbf{What is the most important festival?} The Weaving of the New Bridge (spring). The silk is harvested from the canyon's spiders. The thread is spun by the wardens. The bridge is woven in a single night. The bridge is named after the year. The year is named after a promise. The promise is always the same: the valley will hold.

\textbf{What is the secret that could destroy a powerful person?} The river's rust color is not iron. It is blood---the blood of the serpent that sleeps beneath the canyon. The serpent is wounded. The wound has been bleeding for nine centuries. The Lethai-ar know. The Lethai-ar do nothing. The serpent's dream is the valley's law. If the serpent wakes, the law ends.
\end{almanac}

\newpage
\input{../../generators/crimson-valley.tex}
\newpage