
\part{The Amaranthine Sea \& Northern Shores}

% After \part{The Amaranthine Sea & Northern Shores}
\begin{quote}
  \small
  \textit{``The sea keeps a better ledger than any scribe. The reef remembers every secret whispered above its depths. Do not ask which flag flies highest -- ask which hull leaks slowest.''}
  \noindent --- Dusana of the Raven Road
  \end{quote}

% ------------------------------
% CHAPTER: KAHFAGIA
% ------------------------------
\chapter{Kahfagia}
\emph{The Flag of Storms.}

\noindent
{\Large
\textit{I was born on the tide, which is to say I was born owing. My mother was a net-mender who lost her fingers to a reef-cutter's debt. My father was a bell---that is to say, he rang when the Admiralty pulled his string. I learned to read the lantern codes before I learned my letters, and I learned that every light is a lie waiting to be told.}

\vspace{0.5em}

The Admiralty runs this place like a pirate charter with a budget---and oh, how they love pretending it's legitimate. They control the beacons that lure ships onto reefs, the booms that shatter hulls for salvage rights, and the tariffs that fund their champagne baths. But here's the secret the commodores don't want you to know: the lanterns are afraid of the dark too.

\vspace{0.5em}

I've seen a Mirror-Keeper weep when her reflection showed her a daughter she'd sold for a pilot's license. I've watched a Privateer Captain kiss a silver coin before a raid---the same coin her mother kissed before she never came back. This is not a city of heroes. It is a city of people who learned that the sea is cheaper than mercy.

\vspace{0.5em}

If you want to move cargo through the straits without becoming fish food, get a convoy letter stamped with the Admiral's seal. Or get smart. Or get gone.

\vspace{1em}

\noindent\textit{--- Captain Mira ``the Lucky,'' Sidhi exile, pirate by convenience}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salt fish porridge---dried cod soaked overnight, boiled with hardtack and a splash of lamp oil. The oil is for calories. The fish is for memory. Every Kahfagian child learns to taste the salt and know which coast their grandmother came from.

\textbf{What does a child learn before they can walk?} To count ships by their lanterns. Three lanterns means a Kahfagian merchant---safe. Four means an Oshiiran grain barge---neutral. One, low in the water, means a privateer. Run.

\textbf{What is the first thing you notice about a stranger?} Their hands. Dock-workers have rope-burned palms. Merchants have ink-stained fingers. Privateers have no fingernails---they bite them off in storms.

\textbf{What is the most common crime?} False manifests. Every third ship carries cargo that doesn't appear on its papers. The Admiralty knows. The Admiralty takes a cut. The Admiralty calls it a ''tariff surcharge.``

\textbf{What does no one talk about but everyone knows?} The Mirror-Keepers are not born---they are made. And the making requires a death. A sister, a daughter, a lover. Someone who drowned so the keeper could see through the fog.

\textbf{Where do people go when they die?} The tide takes them. Kahfagia has no cemeteries above ground. The drowned are weighted with chain and dropped in the deep channel. Their families light candles in paper boats and float them out at dusk. If the candle reaches the horizon, the soul is forgiven. If it sinks, the debt is not yet paid.

\textbf{How do you apologize for a serious wrong?} You offer your bell. Every Kahfagian carries a small brass bell on a leather cord---the sound of their name. Offering it to the wronged party means ''I give you my voice. Speak for me.`` Refusing to return the bell after an apology is a declaration of blood feud.

\textbf{What is the most important festival?} The Beacon Lighting (midwinter). Every lighthouse keeper lights a new lamp from the Admiral's flame. If your beacon catches first, your year will be lucky. If it sputters, you will owe a death to the sea.

\textbf{What is the secret that could destroy a powerful person?} The Admiral's lantern code has a ninth signal---never used, never spoken, burned into every keeper's memory. It means \textit{''Kill the witnesses. Sink the evidence. The sea will hide the rest.``}
\end{almanac}

\newpage
\input{../../generators/kahfagia.tex}
\newpage

% ------------------------------
% CHAPTER: ECKTORIA
% ------------------------------
\chapter{Ecktoria}
\emph{Marble \& Fire.}

\noindent
{\Large
\textit{My grandfather was a senator. He voted to burn the Ykrul tribute-pact, and then he voted to burn the inquiry into why the grain ships sank. He signed his name in wax and watched the city burn three times. He never raised his voice.}

\vspace{0.5em}

I am a Coin-House Factor. I seal deals with three ribbons---white for ''ask,`` red for ''now,`` black for ''never.`` I have never used the black ribbon. Not because I am kind, but because ''never`` is a debt I cannot collect.

\vspace{0.5em}

Ecktoria is an Empire that never quite died---more like a corpse propped up by gilded scaffolding and the stubborn refusal of its undertakers to admit the job's done. The Everflame burns in every basilica not as a symbol of divine truth, but as a literal lie detector: swear a false oath near it, and your tongue turns black. I have seen a merchant's tongue turn black mid-sentence. He finished his lie anyway. The buyers pretended not to notice. \textit{That is Ecktoria.}

\vspace{0.5em}

The statues watch. The fire judges. The marble remembers every name carved into its steps---and every name left out.

\vspace{0.5em}

Do not refuse a toast in a coin-house. The ledgers are the true judges, and a declined cup is a declined credit. You will not borrow again.

\vspace{1em}

\noindent\textit{--- Elijah of the Red Ledger, Coin-House Factor, third generation}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Spiced flatbread dipped in olive oil and honey---but only if you can afford the honey. The poor eat millet porridge with a single dried fig. The fig is counted as part of the household's grain allowance.

\textbf{What does a child learn before they can walk?} To count coins by touch. Every Ecktorian child has a box of worn sesterces; by age five, they can tell a debased copper from pure silver by the weight alone.

\textbf{What is the first thing you notice about a stranger?} Their signature. Ecktorians judge by handwriting before face. A clerk with a steady hand is trustworthy. A noble with a flourish is hiding something. A merchant who signs with an X is not worth doing business with.

\textbf{What is the most common crime?} Tax evasion---but it is not a crime, it is a sport. Every household keeps two ledgers: the white one for the Censor, the red one for the family. The art is in keeping the red ledger from looking like a confession.

\textbf{What does no one talk about but everyone knows?} The Censor's clerks are paid to find discrepancies. They are also paid to lose them. The price of a lost discrepancy is exactly 5\% of the tax owed. This is not corruption. It is ''administrative flexibility.``

\textbf{Where do people go when they die?} The family crypt, beneath the basilica floor. The wealthy are burned, their ashes mixed with marble dust and pressed into the floor tiles. The poor are buried in unmarked graves outside the walls. The Everflame's light does not reach that far.

\textbf{How do you apologize for a serious wrong?} You petition the Censor to have the wrong stricken from the public record. The petition costs a tenth of your annual income. If accepted, the wrong is erased---legally, it never happened. If rejected, the wrong is published, and you lose your civic standing.

\textbf{What is the most important festival?} The Triumph of the Flame (autumn equinox). The basilica's central flame is carried in procession through every district. To touch the flame is to be absolved of petty debts. To be burned by it is a sign that your debts are not petty.

\textbf{What is the secret that could destroy a powerful person?} The Grand Magistrate's seal is a forgery. The original cracked in 783 AR. The current seal was carved by a prison engraver who was executed to keep the secret. The engraver's apprentice is still alive. They live in Silkstrand. They have the original sketches.
\end{almanac}

\newpage
\input{../../generators/ecktoria.tex}
\newpage

% ------------------------------
% CHAPTER: ACASIA
% ------------------------------
\chapter{Acasia}
\emph{The Broken Marches.}

\noindent
{\Large
\textit{We do not speak of the Curse. We speak around it, under it, through the gaps where words used to fit. The Curse is not a thing. It is the space where a thing used to be.}

\vspace{0.5em}

I am a charcoal-burner. My family has worked these woods for three generations. My grandfather could name every tree between the Pale Causeway and the Wolfstairs Pass. My father could read a deer's track in packed clay. I can tell you where the villages used to be. The map in my head has more empty spaces than marks.

\vspace{0.5em}

Acasia was not always like this. Once, the granaries held grain through winter. Once, the roads carried more than ghosts and tax collectors. Then the trees started dying. Not all at once---slowly, branch by branch, forest by forest. The Blackwood turned gray, then brittle, then silent. The farmers said it was a curse. The duke said it was blight. The merchants said it was the tariffs. The priests said it was the will of gods who had stopped listening.

\vspace{0.5em}

Nobody said it was guilt. Nobody had to. Guilt is the smell of mildew on old parchment. Guilt is the way a village watches you pass without opening its gates. Guilt is the empty chair at the feast, the unmarked grave behind the chapel, the child who will not meet your eyes because their parents told them not to speak to strangers.

\vspace{0.5em}

The truth---the one we do not speak---is that we could have saved the Blackwood. We saw the first trees wither. We saw the sap run black. We did nothing. We told ourselves it would pass. We told ourselves the duke would fix it. We told ourselves the forest was not our responsibility. And then the forest died, and the silence that followed was our answer.

\vspace{0.5em}

The Curse of Acasia is not a spell. It is a memory. The land remembers what we did---and what we failed to do. Every broken milestone on the old Imperial Road is a forgotten promise. Every abandoned vineyard is a harvest we chose not to plant. Every bridge that leads back to the same shore is the road we refused to take.

\vspace{0.5em}

Never sign a contract on a bridge. The river is a witness, and it never forgets a signature. But more than that: the bridge knows when you are lying. Not to the other party. To yourself.

\vspace{0.5em}

We are the Curse now. We carry it in our silence, in our empty fields, in the way we turn away when someone asks where the children went. Acasia is not a broken province. It is a province that broke itself, one small refusal at a time, until there was nothing left to break but the people who stayed.

\vspace{0.5em}

And still we stay. Because leaving would mean admitting we were wrong. And that is the one thing guilt cannot bear.

\vspace{1em}

\noindent\textit{--- Corvin, charcoal-burner of the Blackwood, who no longer names the trees because there are too few left to name}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Hardtack soaked in broth---if you have broth. If not, hardtack and hope. The granaries have been empty for three winters. The merchants sell grain at famine prices. The duke says the roads are too dangerous for caravans. The farmers say the duke owns the caravans.

\textbf{What does a child learn before they can walk?} To name the dead. Every Acasian child has a list of ancestors they repeat at bedtime. The list gets shorter each generation. Not because children forget. Because the dead are forgetting themselves.

\textbf{What is the first thing you notice about a stranger?} Whether they have a shadow. The Curse takes shadows first. A man without a shadow is not cursed---he is already gone. His body just hasn't stopped moving yet.

\textbf{What is the most common crime?} Bridge-toll evasion. The roads are unsafe, the ferries are unreliable, but the barons still demand payment for crossings no one can make. The crime is not in refusing to pay. The crime is in being caught trying.

\textbf{What does no one talk about but everyone knows?} The Curse began with a promise broken on the Blackwood Bridge. A baron promised his daughter to a lord of the old forest. He married her to a lowland knight instead. The forest remembered. The forest is still waiting for the wedding.

\textbf{Where do people go when they die?} Into the Blackwood. Burial is forbidden---the ground rejects the dead. Bodies are carried to the tree-line and left. The wood takes them. Sometimes, at dusk, you can see them walking between the trunks. They do not look back.

\textbf{How do you apologize for a serious wrong?} You walk the Broken Mile---a stretch of road that collapsed into the river a decade ago. If you reach the far side, you are forgiven. If you fall, you were not sorry enough. No one has reached the far side in seven years.

\textbf{What is the most important festival?} The Counting of the Names (midwinter). Every village lists its dead aloud, from oldest to newest. If a name is forgotten, the ghost walks until someone remembers. The list grows longer each year. The memory grows shorter.

\textbf{What is the secret that could destroy a powerful person?} The duke's son is a ghoul. He was lost in the Blackwood as a child. They found him three days later, pale and smiling. He does not age. He does not eat. He only watches. The servants know. The servants do not speak.
\end{almanac}

\newpage
\input{../../generators/acasia.tex}
\newpage

% ------------------------------
% CHAPTER: VHASIA
% ------------------------------
\chapter{Vhasia}
\emph{The Fractured Sun.}

\noindent
{\Large
\textit{Two crowns. One kingdom. A river so wide it might as well be a border---and oh, how the crowns love pretending it's not.}

\vspace{0.5em}

I am a Road Warden. I have served under three different banners on this same stretch of road, and the only constant is that someone always wants to collect tolls. The Northern Seigneurie and the Southern Parlement both claim the throne, both flood the river with paperwork to ''prove`` their right, and both are too busy arguing over seal wax colors to notice that the Fenwoods already burned Lence, and the Sun-court is a memory.

\vspace{0.5em}

The Last Dauphin? A rumor. But rumors pay well when you're the one who started them---and I started three: that he was swapped at birth with a goatherd's child, that he's secretly a Veiled Sister in disguise, and that he's been dead for ten years, his ''reappearance`` a puppet show run by the Seigneurie's spymaster.

\vspace{0.5em}

Never carry a coin that bears only one sunburst. The realm has three mints, and the wrong metal on the wrong road is treason.

\vspace{1em}

\noindent\textit{--- Road Warden Cerran, who has not slept in two days and is not sure which crown he serves}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Dark bread and watered wine---the wine cuts the dust from the road. The nobility eat spiced eggs and fresh cheese, but they eat behind closed doors. To eat well in public is to invite envy. Envy is a dagger.

\textbf{What does a child learn before they can walk?} Which colors not to wear. Black and gold are the Northern Crown's colors. Red and silver are the Southern Parlement's. A child who wears the wrong color to the wrong village may not return home.

\textbf{What is the first thing you notice about a stranger?} Which side of the road they walk on. The north side is the Crown's. The south side is the Parlement's. The middle is no man's land---literally. Bandits, spies, and messengers use the middle. So do ghosts.

\textbf{What is the most common crime?} Raiding across the Belworth. The river is the border, the border is a suggestion, and the suggestion is enforced by armed men on both sides. The crime is not in raiding. The crime is in being caught on the wrong bank at dawn.

\textbf{What does no one talk about but everyone knows?} The Sun-court was not burned. It was bankrupted. The Fenwoods did not sack Lence---they bought it. Every judge, every clerk, every guard captain had a price. The Fenwoods paid it. The Vhasians are still paying interest.

\textbf{Where do people go when they die?} To the crypts beneath the Sun-court---if they can afford a plaque. The poor are buried in unmarked graves along the Belworth floodplain. The river floods every spring. The dead wash downstream, into the land of the other crown. Neither side claims them.

\textbf{How do you apologize for a serious wrong?} You offer a chevauchée---a raiding party that burns a village in your name. The apology is not in the burning. The apology is in the promise that the village will be rebuilt with your own hands. If you burn a village and do not rebuild, you have not apologized. You have started a feud.

\textbf{What is the most important festival?} The Sun-return (spring equinox). Both crowns claim the sun as their symbol. On this day, both hold processions. The processions are identical. The participants do not acknowledge each other. The sun does not care.

\textbf{What is the secret that could destroy a powerful person?} The Queen-Mother has no claim to the regency. Her son is not the Dauphin. The real Dauphin died in childbirth. The Queen-Mother's daughter was stillborn. The child on the throne is a foundling, bought from a Tulkani caravan. The Tulkani know. The Tulkani are still waiting to be paid.
\end{almanac}

\newpage
\input{../../generators/vhasia.tex}
\newpage

% ------------------------------
% CHAPTER: VITERRA
% ------------------------------
\chapter{Viterra}
\emph{The Hedge-Law Realm.}

\noindent
{\Large
\textit{Every hedge here marks not just a boundary but a principle---the careful balance between order and freedom, justice and mercy. Our law grows like the hedges themselves, tended with precision and patience.}

\vspace{0.5em}

I am a Queen's Justiciar. I have never lost a legal duel. My opponents have a habit of falling down stairs. I do not push them. The stairs \textit{remember their debts.}

\vspace{0.5em}

Viterra breathes with the careful rhythm of precedent and petition. Once a proud kingdom before becoming a Utaran province, it is now reborn under a Warrior Queen who rules by legal precision as much as military might. The realm's true sinews lie not in steel but in writs, charters, and the ancient rights that bind lord to tenant, crown to county, and church to common.

\vspace{0.5em}

But do not mistake us for gentle. The Queen's Peace is not a promise---it is a contract, renewed each morning by every magistrate, every justicar, every clerk who opens a ledger. And across the Belworth, the Vhasians still claim our hedges are theirs. Let them try.

\vspace{0.5em}

Never cross the Belworth without a charter recognized on both banks. Viterran law and Vhasian claim fight every crossing, and the river takes the losers.

\vspace{1em}

\noindent\textit{--- Justiciar Elodie Venn, who invoked the Right of Acknowledgment and watched a father speak truth to his son for the first time}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Barley porridge with a drizzle of honey---if you own bees. If not, porridge with salt. Salt is the currency of the poor. Honey is the currency of the hopeful.

\textbf{What does a child learn before they can walk?} To recite their lineage back three generations. Every Viterran child knows their grandparents' names, their grandparents' trades, and their grandparents' debts. A child who cannot recite their lineage is a child without standing. Standing is everything.

\textbf{What is the first thing you notice about a stranger?} Whether they carry a seal ring. A seal ring means they have the right to make contracts in their own name. A stranger without a seal ring is either a servant or a criminal. In Viterra, the difference is a matter of paperwork.

\textbf{What is the most common crime?} Boundary disputes. A hedge moved three feet east. A milestone replaced under cover of darkness. A ford declared ''private`` when it has been public for two centuries. The courts are clogged with such cases. The lawyers are rich. The hedges are patient.

\textbf{What does no one talk about but everyone knows?} The Queen is not a warrior. She has never drawn a sword. Her reputation was forged by a propaganda campaign run by her spymaster. The spymaster is dead. His successor is more discreet.

\textbf{Where do people go when they die?} To the family plot---if they have a family. The landless poor are buried in parish cemeteries, their headstones marked with a single word: ''Anonymous.`` The ghosts of the anonymous wander the hedgerows at dusk. They do not remember their names. They remember their debts.

\textbf{How do you apologize for a serious wrong?} You petition the hedge-court for a ''writ of contrition.`` The writ requires you to state your wrong aloud, in public, while holding a handful of earth from the wronged party's land. If the earth crumbles, you are forgiven. If it holds, your apology is not accepted. You may try again next season.

\textbf{What is the most important festival?} The Opening of the Rolls (autumn). Every county sends its tax records to the capital. The procession is a week long. The documents are carried by children, so that the next generation may learn the weight of obligation.

\textbf{What is the secret that could destroy a powerful person?} The Codex of Precedents has a missing page. It was torn out fifty years ago, during a trial for treason. The page contained the precedent for executing a monarch. The Queen knows. The Justiciars know. The page is hidden in a lead-lined box beneath the hedge-court at Thornwood. The hedge-court is haunted. The ghost knows where the key is.
\end{almanac}

\newpage
\input{../../generators/viterra.tex}
\newpage

% ------------------------------
% CHAPTER: THEPYRGOS
% ------------------------------
\chapter{Thepyrgos}
\emph{City of a Thousand Stairs.}

\noindent
{\Large
\textit{Every stair climbed is a word in the great conversation between earth and sky. This city is not built---it is grown, stone by stone, law by law, until the very architecture speaks wisdom.}

\vspace{0.5em}

I am a Bell-Runner. I know every step that creaks a warning and every landing that whispers gossip. The bells tell the city's mood better than any lord---three chimes means business, six means trouble, and nine means the hunters are coming.

\vspace{0.5em}

The Synod rules here with the delicacy of a brick wrapped in velvet---they hate unsanctioned magic like it's a personal insult, but they adore infrastructure, especially when it's shiny, expensive, and comes with a side of indulgences they can sell to the faithful. The Chain-Lanterns move at dawn, their brass censers swinging, their bells tuned to the frequency of latent magic. In the university's forbidden tiers, the Witness and the Ninth arm rival fraternities with citations and curses, and no one is sure which side the Archon secretly favors.

\vspace{0.5em}

Never let a bell-ringer count your breaths. The Chain-Lanterns pay well for names that ring false on the test stairs.

\vspace{1em}

\noindent\textit{--- Sofia, Bell-Runner, who ties a knot in her hair for each hunter she has outrun}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Flatbread with olive oil and a boiled egg---if you are a clerk. If you are a student, bread and water. If you are a professor, bread, water, and a single fig. The fig is a status symbol. The status is ''I have not been audited.``

\textbf{What does a child learn before they can walk?} To read a bell schedule. Thepyrgos runs on bells---dawn bells, market bells, curfew bells, trial bells. A child who cannot read the bells cannot cross the city without a guardian. A child without a guardian is a child who will be sold to the university for dissection.

\textbf{What is the first thing you notice about a stranger?} Their step count. Thepyrgos has a thousand stairs. A stranger who counts them aloud is either a pilgrim or a spy. A stranger who climbs them in silence is either a native or a ghost.

\textbf{What is the most common crime?} Unsanctioned magic. The Synod licenses every spell cast within the walls. An unlicensed cantrip is a misdemeanour. An unlicensed healing is a felony. An unlicensed summoning is a death sentence---not for the caster, who will be executed, but for the summoned, who will be bound to the city's foundation as a ''structural consultant.``

\textbf{What does no one talk about but everyone knows?} The Silent Stair is not silent. It whispers. The whispers are the confessions of the dead---scholars who died in the university's service, their secrets recorded in the stone. The Chain-Lanterns listen. They do not share.

\textbf{Where do people go when they die?} To the catacombs beneath the university. The catacombs are a library of bones. Each skeleton is labeled with the scholar's name, their thesis, and the date of their death. The unlabelled bones are the ones who died before they could finish. They are still writing.

\textbf{How do you apologize for a serious wrong?} You submit a ''retraction`` to the Synod's Committee on Ethical Correction. The retraction must be published in the public square, read aloud by a herald, and witnessed by at least three notaries. The fee is 50 silver. The shame is free.

\textbf{What is the most important festival?} The Convocation of the Steps (summer solstice). Every citizen climbs the thousand stairs to the upper citadel. At the top, the Archon reads the names of the dead from the past year. The names are carved into the steps as they are read. The steps are getting crowded.

\textbf{What is the secret that could destroy a powerful person?} The Archon is a fraud. He never passed the Synod's examination. His degree was purchased from a disgraced dean whose signature is forged on every document. The forger is still alive. They live in the catacombs. They have the original quill.
\end{almanac}

\newpage
\input{../../generators/thepyrgos.tex}
\newpage