\part{Dhahara --- The Kingdom of Brass}
% After \part{Dhahara --- The Kingdom of Brass}
\begin{quote}
  \small
  \textit{``The mountain weighs every breath, every ingot, every oath. Pay in iron, pay in silence, or pay in blood. And when the ninth lease cracks, even the elephants will remember who fed them during the famine.''}
  \noindent --- Dusana of the Raven Road, at the Viceroy's Hall in Sarvistan
  \end{quote}
% ------------------------------
% CHAPTER: THE BRASS COAST (REVISED)
% ------------------------------
\chapter{The Brass Coast}
\emph{Shore of Silver and Spice.}

\noindent
{\Large
\textit{''The tariff is the law. The law is the tariff. I do not care if your goods are stolen, blessed, or weeping. I care that you pay. The Admiralty does not rule this coast. The ledger rules. I am merely its pen.``}

\vspace{0.5em}

I am a spice merchant, grandson of a raja who watched the first Kahfagian customs house rise on our beach. The Admiralty calls me a subject. The Sidhi call me a partner. I call myself a son of the Brass Coast. My grandfather's sword hangs above my desk. One day, I may need it again. The sword has nine notches. Each notch is a promise the Admiralty broke.

\vspace{0.5em}

The Brass Coast is Dhahara's southern rim: a humid, feverish coastline of palm groves, salt pans, and crowded port cities. The Kahfagian Admiralty administers it as a ''client province,`` but three powers bleed into each other here, and none can draw a clean breath. The Admiralty holds the ports and the tariff ledgers. The Sidhi diaspora runs the ropeways, the lighthouse courts, and the freedman's seal. And the native Dhaharan rajas --- the hill lords who have lived here since before any foreign flag flew --- grow the pepper, mine the iron, and watch the coast from inland fortresses, waiting.

\vspace{0.5em}

Here, a man can buy anything --- a ship, a spice lot, a freedman's seal --- and sell anything except his own shadow. The Admiralty's Maritime Court settles disputes with a judge's gavel and a marine's cutlass. The Sidhi lighthouse courts offer an alternative: justice by green flame, witnessed by the drowned. And the rajas enforce their own law in the hills, where the Admiralty's writ does not run. But here's the secret the Admiralty doesn''t want you to know: the ropeways that carry spice baskets from ship to warehouse are old, and the dead still pull their weight --- not just Sidhi dead, but the drowned of every nation who could not pay their passage. The coast remembers them all. The ninth rope on every ropeway is always frayed. No one replaces it. That rope carries the debt.

\vspace{0.5em}

\textbf{Never board a ship that has a green lantern at the mast.} It means the captain is under a death vow --- and you are the payment. If the lantern burns blue, run. The blue flame is the Admiral's oracle, and she does not make bargains.

\vspace{1em}

\noindent\textit{--- Raja Govind of the Kalinga foothills, who has never seen the sea and refuses to start}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice and fermented fish paste --- the rice is grown in coastal paddies, the fish is caught by the lighthouse keepers`` children. The children are paid in copper. The copper is melted into seals. The seals are sold to whoever can afford them. The ninth child never takes payment. No one knows why.

\textbf{What does a child learn before they can walk?} To read a lighthouse code. Brass Coast children are taught the eight lamp-flashes: Tariff, Inspection, Quarantine, Smuggler, Pirate, Rescue, Home, and the ninth --- the flash that means ''the collector is coming.`` The ninth flash is never used. The ninth flash is a warning. The children learn it anyway.

\textbf{What is the first thing you notice about a stranger?} Their papers. The Admiralty requires a manifest for every shipment, a seal for every trader, a stamp for every dock. A stranger whose papers are in order is a merchant. A stranger whose papers are missing is a smuggler. A stranger whose papers are too perfect is a spy. The spies are Admiralty. The Admiralty spies on itself. The ninth page of every manifest is always blank. That is where the dead travel.

\textbf{What is the most common crime?} Tariff evasion --- but tariff evasion is not a crime, it is a profession. The tariffs are set by the Admiralty. The Admiralty is corrupt. The evaders pay the Admiralty directly. The payment is called a ''consulting fee.`` The fee is tax-deductible. The ninth fee is always paid in salt, not silver. The salt comes from the Freedman's Grave.

\textbf{What does no one talk about but everyone knows?} The lighthouse lamps are not fuelled by oil. They are fuelled by the fat of the drowned --- whales, sailors, slaves who did not survive the crossing. The fat burns green. The green is the color of memory. The memory is the dead''s. The ninth lamp on every tower is never lit. It is reserved for the collector.

\textbf{Where do people go when they die?} To the sea. The dead are weighted with brass and dropped in the deep channel. The channel is called the Freedman's Grave, but it holds rajas and pirates alike. The grave is unmarked. The grave is the entire coast. The ninth body dropped each year is never weighted. It floats. It always drifts toward the Admiral's palace.

\textbf{How do you apologize for a serious wrong?} You offer a tariff --- a payment of silver, calculated by the wronged party's estimate of their loss. The tariff is paid to the Maritime Court. The court takes a percentage. The percentage is the judge's fee. The judge's fee is negotiable. The ninth tariff ever collected was paid in a memory. The judge still remembers it. He wishes he did not.

\textbf{What is the most important festival?} The Burning of the Ledgers (new year). Every merchant's ledger is brought to the lighthouse. The ledgers are read aloud. The debts are tallied. The ledgers are burned. The ash falls on the water. The water carries the debts to the sea. The sea does not forgive. The sea reschedules. The ninth ledger is never burned. It is kept in a brass box at the bottom of the Freedman's Grave. It contains the Admiralty's own debts.

\textbf{What is the secret that could destroy a powerful person?} The lighthouse lamps are not all green. One lamp burns blue --- the Admiral's personal beacon. The blue flame is not fat. It is a spirit --- a Sidhi oracle who was bound to the lighthouse nine generations ago. She sees everything. She speaks to no one. The Admiral does not know she is still alive. Her ninth prophecy was never recorded. It said: ``The collector will come by sea, and the ledger will burn from the bottom up.''
\end{almanac}

\newpage
\input{../generators/brass_coast.tex}
\newpage

% ------------------------------
% CHAPTER: THE NINE KINGDOMS (REVISED)
% ------------------------------
\chapter{The Nine Kingdoms}
\emph{The Peacock Throne and Its Rivals.}

\noindent
{\Large
\textit{``The Peacock Throne is not a seat. It is a promise made by the first sun king: that the nine kingdoms would remain nine, never one, never fewer. I do not rule. I balance. A peacock's tail has nine eyes. Each eye watches a different kingdom. When one closes, the throne weeps.''}

\vspace{0.5em}

I am a Temple Dancer. My dance is not for the king. It is for the monsoon. If my ankle bells ring true, the rains come. If they falter, the wells dry. They have never faltered. I do not sleep before a performance. I dream the rhythm. And I dream of the missing ninth.

\vspace{0.5em}

Between the Satrapies plateau and the Brass Coast lies a fertile, contested lowland: the land of the Nine Kingdoms. This is the heart of Dhahara, where ancient dynasties still claim descent from the sun, and where the Peacock Throne of Ahimsa casts its shadow over half a dozen rival courts. Here, rivers are borders, elephants are war machines, and every harvest is a negotiation between kings who have never met but whose grandfathers swore blood oaths.

But here's the secret the Maharaja doesn't want you to know: the number ``nine'' is a lie. There have never been nine kingdoms. The number is a ritual --- a prayer that the Hollow will not notice that one of the nine is always missing. The throne weeps because the missing kingdom is always the one that mattered. Its debt is still compounding.

\vspace{0.5em}

The Covenant has a strong presence in the sacred cities; the Aeler have counting houses in the iron towns; the Sidhi sail up the rivers to trade. But the Nine Kingdoms belong to the Dhaharan --- to their poets, their temple dancers, their bandit kings, and their queens who rule from behind silk screens. A king's authority rests on his genealogy --- how many generations back to the sun, how many treaties sealed, how many marriages made. But a king's true power rests on how much of his family's debt has been paid. The Peacock Throne of Ahimsa is the nominal center, but the other eight kings vie for influence, and the balance shifts with every monsoon.

\vspace{0.5em}

\textbf{Never praise a king's elephant.} The elephant is sacred; your praise will be taken as an insult to the king's own mount. Compliment the mahout instead. The mahout will tell the elephant. The ninth elephant in the royal stables has no mahout. It waits.

\vspace{1em}

\noindent\textit{--- Temple Dancer Mira, who has stopped three droughts and started none}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with mango pickle --- the rice is from the royal granaries, the mangoes are from the king's orchards. The king's orchards are irrigated by the river. The river is contested. The contest is a negotiation. The negotiation is breakfast. The ninth grain of rice in every bowl is left uneaten. It is an offering to the missing kingdom.

\textbf{What does a child learn before they can walk?} To recite their genealogy. Nine Kingdoms children are taught the names of their ancestors back to the sun. A child who cannot recite is a child without a kingdom. A child without a kingdom is a servant. Servants are not fed. The ninth ancestor is never named. That is the gap where the Hollow sits.

\textbf{What is the first thing you notice about a stranger?} Their caste mark. The Nine Kingdoms have nine castes --- priest, warrior, merchant, farmer, artisan, servant, outcast, foreigner, and the ninth --- the caste that has no name. The ninth caste is the king's spies. The spies are everywhere. The spies have no caste mark. Their faces are always the same. No one can describe them.

\textbf{What is the most common crime?} Tax evasion --- but tax evasion is not a crime, it is a political statement. The king's tax collectors are also the king's spies. A merchant who evades taxes is a merchant who has something to hide. The king wants to know what. The king will find out. The ninth tax collector vanished last year. His shadow still walks the counting house.

\textbf{What does no one talk about but everyone knows?} The Peacock Throne is empty. The Maharaja died seven years ago. His death was concealed by the courtiers, who rule in his name. The courtiers are the nine ministers. The ninth minister is a eunuch who controls the palace guards. The eunuch has the real power. He has no shadow. The guards do not notice.

\textbf{Where do people go when they die?} To the river. The dead are cremated on the ghats. Their ashes are scattered on the water. The water flows to the sea. The sea is the kingdom's border. The border is a negotiation. The negotiation never ends. The ninth ghat is reserved for the missing kingdom. It has never been used. It is always warm.

\textbf{How do you apologize for a serious wrong?} You offer a dance --- a performance at the temple of the sun. The dance is judged by the priest. If the priest smiles, you are forgiven. If the priest weeps, you are not. If the priest is silent, you are cursed. The curse is the sun's glare. The sun will not look away. The ninth dance ever performed was for a debt to the Hollow. The dancer is still dancing. No one watches.

\textbf{What is the most important festival?} The Coronation of the Monsoon (summer). The king offers a golden elephant to the rain clouds. The clouds accept or refuse. The clouds have refused for nine years. The king's treasury is empty. The king's patience is not. The ninth golden elephant was hollow. It was filled with ashes. The ashes belonged to the ninth king.

\textbf{What is the secret that could destroy a powerful person?} The golden elephant is hollow. It is filled with the ashes of the ninth king --- the one who tried to unite the kingdoms and failed. His ghost rides the monsoon winds. His ghost is still trying. His ghost will never succeed. But every year, the clouds refuse because they are waiting for him to ask. He never does.
\end{almanac}

\newpage
\input{../generators/nine_kingdoms.tex}
\newpage

% ------------------------------
% CHAPTER: THE HIMADRI FOOTHILLS
% ------------------------------
\chapter{The Himadri Foothills}
\emph{The Iron Passes (Aeler Focus).}

\noindent
{\Large
\textit{The mountain does not speak in words. It speaks in the crack of a bearing, the echo of a hammer, the taste of ore on the tongue. We Aeler do not conquer stone. We negotiate with it. The lease is not a document. It is the weight of the keystone that holds the roof.}

\vspace{0.5em}

I am a Mine-Breath. I carry a bell that counts my breaths. Eight is safe. Nine is the mountain's. My grandfather counted to nine once. The mountain took his shadow. He lived another thirty years, but he never cast a shadow again.

\vspace{0.5em}

The Ninth Seal mine isn't just holding a demon---it's a seal market, and oh, how the Vent-Prior loves a good deal on demon containment. The Aeler do not rule through armies; they rule through the maintenance contracts that keep the mines from collapsing, the bridges from falling, and the passes from freezing. A vent-prior's authority rests on his ability to read the mountain's cracks; a mine-captain's on her skill with a tuning fork. But here's the secret the Vent-Prior doesn't want you to know: the demon in the Ninth Seal is not the only thing the mountain is holding. There are older things deeper down. The Aeler sealed them first. The Ninth Seal is just the newest.

\vspace{0.5em}

The elephant caravans haul ore down to the Brass Coast; mule trains carry silver and tea up to the mountain holds. The Aeler engineers speak a creole of their own tongue; the Dhaharan nobles wear Aeler breath-counters as jewelry; and deep beneath the oldest mine, a demon sleeps behind a Covenant seal---a fragment of the thing that turned the Red Desert to glass. The seal is cracking. The demon whispers through the cracks. It offers favors. It asks for small betrayals. It is very patient.

\vspace{0.5em}

Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one---a breath, a memory, a shadow.

\vspace{1em}

\noindent\textit{--- Mine-Breath, who carries her grandfather's bell and has never reached nine}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Iron rations and tea---the rations are hardtack and dried meat, the tea is brewed from leaves that grow only above the snow line. The leaves are harvested by the Vent-Prior's apprentices. The apprentices are chosen by lottery. The lottery is a death sentence.

\textbf{What does a child learn before they can walk?} To read a pressure gauge. Himadri children are taught the eight pressures: Safe, Warning, Danger, Collapse, Flood, Fire, Seal, and the ninth---the pressure that means the mountain is breathing. The mountain's breath is the miners' death.

\textbf{What is the first thing you notice about a stranger?} Their bell. Every miner carries a breath-counter bell. A stranger whose bell rings too fast has been running. A stranger whose bell is silent is dead. A stranger whose bell rings the ninth note is a demon. The demons are sealed. The seals are cracking.

\textbf{What is the most common crime?} Ore smuggling---but ore smuggling is not a crime, it is a perquisite. The mine-captains take a cut of every shipment. The cut is called ``the Vent-Prior's fee.'' The fee is negotiable. The negotiation is a ritual. The ritual involves tea and threats.

\textbf{What does no one talk about but everyone knows?} The Ninth Seal is not a seal. It is a door---a gate to a prison that holds something that should not be named. The seal's keepers do not know what is inside. The keepers do not want to know. The keepers are paid to not know.

\textbf{Where do people go when they die?} To the mountain. The dead are not buried. Their bodies are left in worked-out galleries. The mountain absorbs them. The mountain's memory grows. The mountain remembers every breath it has taken.

\textbf{How do you apologize for a serious wrong?} You offer a breath---one of your own, counted and witnessed. The breath is sealed in a brass capsule and placed in the Vent-Prior's ledger. The ledger is the court. The court's judgment is the mountain's. The mountain's judgment is final.

\textbf{What is the most important festival?} The Counting of the Breath (winter solstice). Every miner's breath-counter is read aloud. The miners with the lowest counts are promoted. The miners with the highest counts are retired. Retirement is a sealed gallery. The gallery is a tomb.

\textbf{What is the secret that could destroy a powerful person?} The demon in the Ninth Seal is not a demon. It is an Aeler---the first Vent-Prior, who volunteered to be sealed with the mountain's hunger. He is still alive. He has been counting his breaths for nine centuries. He has reached the ninth count. He is still counting.
\end{almanac}

\newpage
\input{../generators/himadri_foothills.tex}
\newpage