\part{The Eastern Realms}
% After \part{The Eastern Realms}
\begin{quote}
  \small
  \textit{``The maps of the empire draw the East as a single sweep -- Sihai, Nihori, Ayokha -- all pressed together like tiles on a game board. The roads know better. Sihai measures time in dynasties; Nihori counts it in storms. Ayokha negotiates between them with the patience of the monsoon. And beneath them all, the older peoples keep their own thresholds, their own debts, their own ways of not being seen.''}
  \noindent --- Dusana of the Raven Road, at the edge of the Inland Sea
  \end{quote}
% ------------------------------
% CHAPTER: SIHAI
% ------------------------------
\chapter{Sihai}
\emph{The Central Kingdom.}

\noindent
{\Large
\textit{The Mandate of Heaven is not a right. It is a loan. The emperor borrows legitimacy from the ancestors, from the land, from the Hollow itself. When the loan is due, the empire falls. I do not serve the emperor. I serve the ledger that tracks the interest.}

\vspace{0.5em}

I am a Provincial Censor. My brush is heavier than any sword. A single stroke can erase a name, a family, a dynasty. The emperor reads my reports and trembles. He knows that I have written nine reports this year. The ninth is about him.

\vspace{0.5em}

Sihai is the central kingdom, the seat of the Emperor and the Mandate of Heaven. Its canals carry grain from the south; its examination halls produce bureaucrats who govern by the brush. Power here is measured in ink, not iron. The Censorate watches everyone; the Salt Monopoly controls trade; the Lethai-thora keep the secrets that the Mandate prefers to forget. But here's the secret the Emperor doesn't want you to know: the Mandate is not a divine right. It is a loan. The ancestors are the creditors. The Hollow collects the interest. Every dynasty that has ever fallen did so because the interest came due and the emperor could not pay.

\vspace{0.5em}

Beneath the mortal bureaucracy runs a deeper one: the Celestial Bureaucracy of gods, ghosts, and ancestors, where every official has a heavenly counterpart and every erased name becomes a hungry ghost. The Lethai-thora---the ancient archivists who taught the first emperors to write---still keep the hidden ledgers, the ones that record what the Censorate burns. A well-written couplet can appease a ghost; a forged seal can summon a demon. The Hollow's ninth debt is the unpaid interest on the Mandate itself.

\vspace{0.5em}

Never speak of succession before naming the Emperor. The Censorate has ears in every tea house. A misplaced word can erase your name from the rolls.

\vspace{1em}

\noindent\textit{--- Quiet Chancellor Wei, who has never smiled and whose seal is a laughing mouth}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Congee with preserved vegetables---the congee is rice boiled until it is memory, the vegetables are salted and aged in clay jars. The jars are sealed with the emperor's own wax. The wax is a promise. The promise is that the vegetables will not poison you. The promise is sometimes broken.

\textbf{What does a child learn before they can walk?} To write their name. Sihai children are taught the eight strokes of the character ``person''---the ninth stroke is the one that turns a name into a signature. A signature is a contract. A contract is a debt. A debt is a life.

\textbf{What is the first thing you notice about a stranger?} Their chop---the seal they carry to sign documents. A stranger whose chop is jade is a noble. A stranger whose chop is brass is a clerk. A stranger whose chop is wood is a merchant. A stranger with no chop is a ghost. Ghosts cannot sign contracts. Ghosts cannot be trusted.

\textbf{What is the most common crime?} Examination fraud---but fraud is not a crime, it is a family business. The examinations determine who governs. The questions are leaked in advance. The answers are memorized. The examiners are bribed. The system is a sieve. The sieve is called ``merit.''

\textbf{What does no one talk about but everyone knows?} The Emperor is a figurehead. The real power is the Censorate---nine ministers who have never been seen together. Their names are not recorded. Their faces are not painted. Their orders are whispered. The Emperor signs what they tell him to sign. The Emperor has not refused in forty years.

\textbf{Where do people go when they die?} To the ancestral tablets. The dead are not buried. Their names are written on wooden tablets, placed in the family shrine. The tablets are fed with rice and tea. The dead eat the steam. The dead are never full.

\textbf{How do you apologize for a serious wrong?} You offer a calligraphy---a scroll written in your own blood. The scroll is presented to the wronged party's ancestral shrine. If the ancestors accept, you are forgiven. If the scroll bursts into flame, you are not. If the scroll writes itself, the ancestors are angry. The ancestors' anger is a famine.

\textbf{What is the most important festival?} The Festival of the Ancestors (autumn). The tablets are cleaned and polished. The dead are invited to feast. The feast is laid out on the family altar. The food is removed after an hour. The dead have eaten the essence. The essence is memory.

\textbf{What is the secret that could destroy a powerful person?} The Censorate's ninth minister is a woman---the emperor's own mother, who was supposed to have died in childbirth. She faked her death to escape the politics of the court. She has been running the Censorate from a hidden villa in the western hills. The emperor does not know. The emperor's mother signs his death warrants. He signs them too.
\end{almanac}

\newpage
\input{../generators/sihai.tex}
\newpage

% ------------------------------
% CHAPTER: NIHORI
% ------------------------------
\chapter{Nihori}
\emph{The Isles of the Dawn Spirit.}

\noindent
{\Large
\textit{I ruled the waves before the Hollow learned to walk. My empire sank because I spoke a name I should not have. Do not speak my name. Do not ask what it was. The sea remembers. That is enough.}

\vspace{0.5em}

I am a Wanderer. I carry a broken sword and a grudge. The sword was my father's. The grudge is mine. The man who killed him died nine years ago. I still walk the isles, looking for his ghost. The ghost is looking for me too.

\vspace{0.5em}

Nihori is a chain of volcanic islands, shrouded in mist and lashed by storms. Here, the old empire sank beneath the waves, and the Lethai-al guard the drowned kingdom's memory. The people are wanderers, duelists, and shrine maidens. They worship the storm spirits and fear the ninth wave. The Hollow breathes in the salt foam; the Oath-Keepers ride the tide. But here's the secret the Shogun doesn't want you to know: the old empire did not sink. It was pulled under. And whatever pulled it is still down there, waiting for the ninth wave to break the right way.

\vspace{0.5em}

The Shogun sits in a half-ruined castle, his authority challenged by three rival daimyo who sharpen their blades and their treaties. Oni cults whisper in the ash wastes, offering forgotten bargains. The Storm Emperor's ghost does not rule---but his silence haunts every island. The ninth taboo is absolute: never speak the name of the drowned capital. The sea will hear you, and it will ask why you called. And when the sea asks a question, it expects an answer---in blood, in memory, or in a name you did not mean to give.

\vspace{0.5em}

Never speak the name of the drowned capital. The sea will hear you, and it will ask why you called.

\vspace{1em}

\noindent\textit{--- The Storm Emperor's Ghost, who appears as a column of steam over the water}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Grilled fish and pickled vegetables---the fish is caught in the caldera lakes, the vegetables are fermented in clay pots buried in volcanic ash. The ash is warm. The warmth is the mountain's breath. The mountain is sleeping. The mountain dreams of the drowned empire.

\textbf{What does a child learn before they can walk?} To read the tide. Nihori children are taught the eight wave-heights: Calm, Ripple, Swell, Storm, Tsunami, and the ninth---the wave that has no height because it is the sea itself rising. The ninth wave is the drowning. The drowning is the empire's memory.

\textbf{What is the first thing you notice about a stranger?} Their sword. Nihori warriors carry their ancestors' blades, passed down for generations. A stranger whose sword is rusted has forgotten their lineage. A stranger whose sword is sharp has remembered. A stranger with no sword is a coward. Cowards are not fed.

\textbf{What is the most common crime?} Dueling---but dueling is not a crime, it is a conversation. The dueling codes are written in verse. The verses are interpreted by the Shogun's court. The interpretations are contradictory. The contradictions are settled by dueling. The circle is beautiful.

\textbf{What does no one talk about but everyone knows?} The drowned capital is not underwater. It is beneath---beneath the caldera, beneath the bedrock, beneath the world. The sea is a lid. The lid is cracked. Something is leaking through. The leaks are the oni. The oni are the empire's regrets.

\textbf{Where do people go when they die?} To the sea cliffs. The dead are placed in caves overlooking the water. The waves sing to them. The songs are the empire's lullabies. The dead listen. The dead do not sleep.

\textbf{How do you apologize for a serious wrong?} You offer a duel---a matched contest with the wronged party's choice of weapon. If you lose, you are forgiven. If you win, you are not. If you refuse, you are exiled. Exile is the open sea. The sea is patient. The sea will wait.

\textbf{What is the most important festival?} The Calling of the Tides (spring). The shrine maidens walk into the water, singing the names of the drowned. The waves answer. The waves speak the names of the living who will die in the coming year. The names are never forgotten. The names are never repeated.

\textbf{What is the secret that could destroy a powerful person?} The Shogun is a ghost. He died in the last oni war, but his body was never found. The daimyo rule in his name because they are afraid of what will happen if they admit he is gone. The oni know. The oni are waiting for the admission. The admission will be the war.
\end{almanac}

\newpage
\input{../generators/nihori.tex}
\newpage

% ------------------------------
% CHAPTER: AYOKHA
% ------------------------------
\chapter{Ayokha}
\emph{The Monsoon Throne.}

\noindent
{\Large
\textit{The monsoon does not come because we pray. It comes because the stars align. I read the stars. I align them. Without me, the rains would fail, the rice would burn, and the kingdom would fall. I am not a servant of the God-King. I am his monsoon.}

\vspace{0.5em}

I am a River Merchant. My boat has more secrets than cargo. The lower deck is for silk. The upper deck is for lies. The tiller is for bribes. I have carried nine spies this year. The ninth was the God-King's own daughter.

\vspace{0.5em}

Ayokha is the land of the Monsoon Throne, where the God-King rules by the rhythm of the rains. Rivers flood, rice terraces glow green, and the wharves are crowded with ships from Dhahara, Sihai, and distant isles. The Lethai-ar witness every contract; the spirit-mediums speak for the nats of the ford. Power here is measured in monsoon schedules, silk vouchers, and the favor of the God-King's astrologer. But here's the secret the astrologer doesn't want you to know: the monsoon does not come because he aligns the stars. The monsoon comes because the river-spirits allow it. And the river-spirits are angry.

\vspace{0.5em}

The nats---hungry spirits who live in eddies and under bridges---demand offerings of betel leaf, candle wax, and forgotten promises. A merchant who ignores a nat's demand will find his boat sunk and his name struck from the ledgers. A farmer who disrespects the monsoon will find his fields barren and his children dreaming of drowning. In Ayokha, the rain is not weather---it is a creditor. And the God-King is behind on his payments.

\vspace{0.5em}

Never refuse a gift of betel leaf. It is an oath of hospitality. Refusing it is an insult that can end a bloodline.

\vspace{1em}

\noindent\textit{--- The God-King's Astrologer, who has not seen the sun in thirty years}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice congee with dried shrimp---the rice is from the terraces, the shrimp is from the delta nets. The delta is where the river meets the sea. The sea is the God-King's mother. The God-King's mother is a nat. The nat is hungry.

\textbf{What does a child learn before they can walk?} To read the river. Ayokha children are taught the eight currents: Flood, Ebb, Eddy, Whirl, Still, and the ninth---the current that flows upstream. The upstream current is the monsoon's breath. The monsoon's breath is the God-King's voice. The God-King is silent.

\textbf{What is the first thing you notice about a stranger?} Their offering. Every traveler in Ayokha carries a gift for the nats---betel leaf, candle, a coin. A stranger whose offering is generous has been blessed. A stranger whose offering is small is testing the spirits. A stranger with no offering is a fool. The nats take fools first.

\textbf{What is the most common crime?} Nat-bribery---but bribery is not a crime, it is a negotiation. The nats set the tolls. The merchants pay the tolls. The nats spend the tolls on offerings. The offerings are burned. The ash falls on the fields. The fields are fertile. The circle is the economy.

\textbf{What does no one talk about but everyone knows?} The God-King is a puppet. The astrologer is the real ruler. The astrologer reads the stars. The stars say what the astrologer tells them to say. The astrologer has not read the stars in twenty years. He reads the nats' moods instead. The nats are fickle.

\textbf{Where do people go when they die?} To the river. The dead are cremated on the ghats. Their ashes are scattered on the water. The water carries them to the delta. The delta is the land of the nats. The nats decide whether the dead become ancestors or hungry ghosts. The nats are not generous.

\textbf{How do you apologize for a serious wrong?} You offer a nat---a spirit of the river, summoned by a medium. The nat hears your apology. If the nat accepts, the water is calm. If the nat refuses, the water rises. If the nat is silent, the water is still. Still water is the God-King's judgment. The God-King does not forgive.

\textbf{What is the most important festival?} The Welcoming of the Monsoon (summer). The God-King stands in the rain. The rain is the nats' blessing. The blessing is measured in inches. The inches determine the harvest. The harvest determines the taxes. The taxes determine the God-King's wealth. The God-King is not wealthy.

\textbf{What is the secret that could destroy a powerful person?} The God-King's daughter is not his daughter. She is the child of the astrologer and a nat---born from a bargain made during a failed monsoon. She can control the rain. She does not know. The astrologer knows. The astrologer is waiting for the right moment to tell her. The right moment is the next drought.
\end{almanac}

\newpage
\input{../generators/ayokah.tex}
\newpage

\chapter{Aelaerem of the Eastern Realms}
\label{chap:aelaerem-east}

\emph{The Hollow's Neighbors.}

\noindent
{\Large
\textit{My grandmother taught me that the Hollow is not evil. It is a forgetfulness. It forgets where the boundary lies, forgets that we are not food, forgets that it was once something else. So we remind it. With a thread, a bell, a bowl of salt. We are its memory.}

\vspace{0.5em}

I am a Thread-Binder. I tied a red thread around my wrist this morning. The Hollow walked past my door last night. It saw the thread. It kept walking. That is the bargain. I give it a boundary. It respects the boundary. Mostly. But last week it forgot the thread's color. I had to tie two.

\vspace{0.5em}

The Aelaerem of the eastern realms have adapted their hedge-magic to thresholds of a different kind: bamboo groves that whisper in the dark, temple stairs that count your heartbeats, and the silence after a storm bell that has no clapper. Here, the Hollow wears the face of the uncanny --- not hunger, but absence. It is the space between two bamboo stalks, the step that is always missing, the name you cannot quite remember.

They survive through ritual precision: a red thread tied at a crossroads with exactly eight knots, a bowl of rice left at dusk and never checked until dawn, a single bell rung once and then muffled with a cloth that has been washed in salt water. They whisper that the Hollow is the shadow of Kuva, cast when the first sun turned away, and that it wanders the world looking for the name it forgot.

Unlike western Aelaerem, who live in close-knit villages, the Eastern Aelaerem are often solitary or live in small family groups---yamabushi of the boundary, shrine hermits of the between. They are the ones who know where the veil is thinnest, and they are the ones paid to keep it closed. Their magic is humble: a knot to bind a promise, a thread to guide a lost ghost, a cup of sake to buy a night's silence.

\vspace{0.5em}

Never count the steps between your door and the crossroads. The Hollow counts with you. If you reach nine, it will be waiting at the ninth step. It will ask, ``Where are you going?'' Do not answer.

\vspace{1em}

\noindent\textit{--- Hedge-Keeper Mila, who has never slept indoors}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
\textbf{What you notice first:} Red threads tied to every gate, every bridge, every well. The smell of burnt rice and old salt. The way the Aelaerem never whistle after dusk---the Hollow thinks you are calling it. And the way they never, ever count to nine.

\textbf{The one rule that kills newcomers:} Never count the steps between your door and the crossroads. The Hollow counts with you. If you reach nine, it will be waiting at the ninth step. It will ask, ``Where are you going?'' Do not answer.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Unspoken}

Power here is measured in red threads tied, offerings left, and the gaps between words. The Aelaerem have no leaders, no courts, no written laws. Each hedge-keeper knows her own stretch of boundary. She knows where the Hollow walks, when it is restless, and what gesture will turn it aside. Her authority is not given by any king --- it is earned by surviving the uncanny. The Aelaerem are not feared or respected; they are tolerated, because without them the Hollow would seep into every threshold. They are the neighbors of the thing that lives between the notes of a song.

\subsection*{Starting Location}

A crossroads at dusk, where a red thread is tied around a wooden post. An old Aelaerem woman is pouring a cup of sake onto the ground. She looks up: ``The Hollow walked last night. I left salt. It took the salt. It left a single red thread tied around my wrist. It moved. Sleep with that knot, and it will know where you are. Do you want to know where it went? It went toward your camp.''

\subsection*{The Almanac of Ordinary Days (Eastern Aelaerem)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with pickled vegetables---the rice is from the paddies, the vegetables are salted in clay pots buried at the boundary. The salt is blessed by the hedge-keeper. The vegetables are always a little bitter. The bitterness is a reminder.

\textbf{What does a child learn before they can walk?} To tie the eight boundary knots. The ninth knot is never tied---it is left loose for the Hollow to find. If a child ties the ninth knot, the Hollow will visit that night and ask for a name. The child's name.

\textbf{What is the first thing you notice about a stranger?} Their shadow. A stranger whose shadow is too short has been touched by the Hollow. A stranger whose shadow moves independently is already lost. A stranger with no shadow is the Hollow itself, wearing a borrowed face. Do not invite them in.

\textbf{What is the most common crime?} Thread-theft---stealing a red thread from a boundary marker. The thread is not valuable, but cutting it breaks a promise. The Hollow becomes restless. The thief is marked: their shadow will shrink one inch each night until they return the thread.

\textbf{What does no one talk about but everyone knows?} The Hollow is not a single thing. It is a place---the space between heartbeats, the pause before a sneeze, the silence after a bell. The Aelaerem do not guard against a monster. They guard against the gaps.

\textbf{Where do people go when they die?} To the salt. The dead are cremated, and their ashes are mixed with salt and scattered along the boundary. The salt draws the Hollow's attention away from the living. The dead become the new boundary. They do not rest. They watch.

\textbf{How do you apologize for a serious wrong?} You leave a cup of sake and a lock of hair at the place where the wrong was done. The Hollow will witness. If the sake is gone by dawn and the hair is untouched, you are forgiven. If the hair is gone, the Hollow has taken your apology as a promise. You are now bound. Do not break that promise.

\textbf{What is the most important festival?} The Retting of the Threads (spring). All red threads on the boundary are taken down, washed in salt water, and re-tied. The village is silent for one night---no bells, no songs, no whispers. The Hollow listens. The Hollow remembers. If a thread is tied wrong, the Hollow will correct it. The correction is always a death.

\textbf{What is the secret that could destroy a powerful person?} The first red thread was not tied by an Aelaerem. It was tied by the Hollow itself, before it forgot what it was. The thread is still there, at the First Boundary. If it is untied, the Hollow will remember its true name. Then it will stop being hungry. It will become something else. No one knows what.
\end{almanac}

\newpage
\input{../generators/aelaerem_east.tex}
\newpage

\chapter{Aelinnel of the Inland Sea}
\label{chap:aelinnel}

\emph{The Bell-Tenders.}

\noindent
{\Large
\textit{The Hollow is not a place. It is a miscount. The ninth step that is never there, the ninth cup that is never poured, the ninth name that is never spoken. We do not fight the Hollow. We leave one empty chair, one unlit lamp, one uncounted stone. That is the price of sitting at the table.}

\vspace{0.5em}

I asked an Aelinnel bell-tender how she kept the tide-spirits from flooding her village. She tapped her temple. ``I count them,'' she said. ``One, two, three, four, five, six, seven, eight. Then I stop. They know that if I ever count nine, I am not counting them.''

\vspace{0.5em}

The Aelinnel of the East do not hide from the spirits --- they coexist. Small spirits inhabit every rock, tree, and wave. The Aelinnel have learned to negotiate with them using copper bells, precise counting, and occasional offerings of pickled plums. They are the engineers of shrines, the builders of floating bridges, and the keepers of the bell-lines that keep tide-spirits from straying. Some say Kuva, the sky, taught them the first counting song, and that the Hollow is simply the space where the count fails.

Unlike their western kin, who lean toward the Gift of the Mind, the Eastern Aelinnel practice a folk mathematics that blends logic and superstition. They believe every spirit has a number-name --- a frequency that can be rung out of a bell, a sequence of steps that can be walked, a knot that can be tied. To offend a spirit is to miscount.

\vspace{0.5em}

Never count to nine aloud near a shrine. The spirits count with you. When you reach nine, they will ask: ``What comes after nine?'' The answer is always the Hollow.

\vspace{1em}

\noindent\textit{--- Bell-Warden Tumble, who writes letters on copper wax}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
\textbf{What you notice first:} The sound of copper bells --- not ringing in sequence, but chiming in overlapping waves. The way every bridge has a missing ninth plank. The smell of pickled plums and salt.

\textbf{The one rule that kills newcomers:} Never count to nine aloud near a shrine. The spirits count with you. When you reach nine, they will ask: ``What comes after nine?'' The answer is always the Hollow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Count}

Power here is measured in bells rung, tides predicted, and the precision of a threshold. The Aelinnel have no kings or councils; their authority rests on the bell-wardens who maintain the shrines and the tide-counters who read the waves. A village that loses its bell-warden is a village that the tide-spirits will flood. A bridge without its ninth plank missing is a bridge that the Hollow will cross. The Aelinnel do not worship the spirits; they transact with them. An offering is not a prayer --- it is a payment. A bell is not a musical instrument --- it is a contract.

\subsection*{Starting Location}

A floating shrine on the Inland Sea, where copper bells hang from every corner. A Bell-Warden rings a bell eight times, then stops: ``The tide-spirit is restless. It wants a ninth bell. I will not give it. You --- count the waves. Tell me how many break before the sun sets. If you miscount, the spirit will ask you for a story. Do not miscount.''

\subsection*{The Almanac of Ordinary Days (Aelinnel)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with pickled plums --- the plums are fermented in salt brine and stored in copper-lidded barrels. The brine is never poured out; it is passed down through generations. The oldest batch is said to be nine generations old. No one has tasted it.

\textbf{What does a child learn before they can walk?} To count the eight safe numbers. Children are taught to count pebbles, footsteps, and bell-rings --- but always to stop at eight. The ninth is never spoken. A child who says ``nine'' aloud is sent to sit by the missing plank of the bridge until they forget they ever said it.

\textbf{What is the first thing you notice about a stranger?} Their fingers. Aelinnel tide-counters often have only eight fingers --- the ninth was a miscount they paid for. A stranger with nine fingers has never made a mistake. A stranger with seven has paid a heavy debt. A stranger with no fingers is either a spirit or someone who counted to nine and lost everything.

\textbf{What is the most common crime?} Bell-theft --- but not for the copper. A stolen bell can be rung at the wrong time, creating a miscount. The thief is not punished by law; they are given to the tide-spirit whose count was broken. The spirit decides the price.

\textbf{What does no one talk about but everyone knows?} The First Bell was not cast by an Aelinnel. It was found --- already ringing, already counting. No one knows who made it. No one knows why it rang. The shards are scattered across the Inland Sea. When the ninth shard is found, the First Bell will ring again. The Hollow will hear it.

\textbf{Where do people go when they die?} To the bell-line. The dead are not buried. Their ashes are mixed with copper dust and cast into new bells. The bells are hung on the bell-lines. When the tide-spirits ring them, the dead speak. Mostly they say: ``Count carefully.''

\textbf{How do you apologize for a serious wrong?} You offer a count --- a public recitation of the wrong, spoken in eights. ``I took the offering. I broke the bell. I counted to nine.'' Then you wait. If a spirit rings a bell eight times, you are forgiven. If it rings nine times, you are cursed. If no bell rings, the Hollow will visit you at midnight.

\textbf{What is the most important festival?} The Retuning of the Bells (autumn). Every bell on the Inland Sea is struck once, in sequence, from the highest tide to the lowest. The sound travels for miles. The spirits listen. If a bell is out of tune, the spirits will correct it by morning. The correction is always a death --- but the bell will ring true for another year.

\textbf{What is the secret that could destroy a powerful person?} The ninth bell was never meant to be cast. It was a mistake --- a foundry worker poured one bell too many. They tried to hide it, but it rang once on its own. The tone was the Hollow's name. The bell still exists. It is buried under a shrine on a small island. No one knows which island. The map was miscounted.
\end{almanac}

\newpage
\input{../generators/aelinnel_inland_sea.tex}
\newpage

\chapter{Lethai-al of Nihori}
\label{chap:lethai-al}

\emph{The Storm-Touched Wanderers.}

\noindent
{\Large
\textit{We do not own the land. We walk it. The road is our shrine. The storm is our law. The sea does not forgive, but it does forget. If you can wait long enough, the waves will erase your debt. But we do not wait. We walk. From shrine to shrine, from storm to storm, we carry the names of the drowned. The road is a prayer. Every step is a syllable.}

\vspace{0.5em}

My grandmother drowned in the ninth wave. Her name became a tide-charm. I keep it in a pouch of oiled leather. When the storm grows angry, I whisper her name. The sea remembers her. It calms for a moment --- long enough to breathe. We are wanderers. The road is our home. The dead travel with us.

\vspace{0.5em}

The Lethai-al of Nihori once guarded a drowned kingdom --- an archipelago that sank when a forbidden name was spoken. Their blood still carries the salt of that ancient catastrophe. Now they are wanderers in the oldest sense: masterless, landless, bound to no lord but fiercely bound to each other and to the storm spirits that swirl above the waves. They walk from shrine to shrine, from port to port, following the processional roads that circle the island.

They do not farm. They do not build. They carry everything they own on their backs or in wooden carts. Their duels are not to the death --- they are to the first draw of blood, after which the loser must join the procession and carry the winner's burden for a season. Their courts are crossroads. Their temples are the road itself. The Eastern Ninth lives here: the Hollow does not hide under hills --- it breathes in the salt foam. The ninth wave is never the largest; it is the one that remembers your name.

\vspace{0.5em}

Never walk under a torii gate without making an offering. The gate is a threshold. The road on the other side is not the same road you left.

\vspace{1em}

\noindent\textit{--- Procession Elder Kenji, who has walked the circuit 99 times}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
\textbf{What you notice first:} The smell of salt and incense. The way the Lethai-al never sit with their backs to the sea. The constant, low murmur of waves being predicted --- not by instruments, but by the tilt of a head, the twitch of a finger. The worn wooden plaques hung from their belts, each carved with a dead ancestor's name.

\textbf{The one rule that kills newcomers:} Never walk under a torii gate without making an offering. The gate is a threshold. The road on the other side is not the same road you left.
\end{tcolorbox}

\subsection*{Regional Mood: The Processional Road}

Power here is measured in shrines visited, storms survived, and the number of dead ancestors carried as tide-charms. The Lethai-al have no king, no courts, no written law. Their authority is personal and processional: a Procession Elder's word is law because she has walked the circuit more times than anyone else. Their guilds are not guilds --- they are processional bands, each with its own rhythm, its own counting song, its own patron shrine. A band that walks too fast offends the spirits; one that walks too slow loses the tide. The Lethai-al do not believe in ownership of land; they believe in passage rights, storm oaths, and the memory of the drowned. Every crossroads is a court. Every storm is a messenger. Every sunrise is a departure.

\subsection*{Starting Location}

A procession trail winding along the sea cliffs, marked by weathered torii gates and small stone shrines. A group of Lethai-al walk in single file, chanting a counting song. One carries a palanquin with a covered object --- a tide-charm shrine. The leader stops: ``You stand on the processional road. The drowned are watching from the waves. Do you walk with us? If so, you must carry nothing. The road provides. The road takes away.''

\subsection*{The Almanac of Ordinary Days (Lethai-al)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Dried fish and rice balls --- the fish is smoked over driftwood fires, the rice is pressed into shapes that fit in a pouch. The food is always cold. The Lethai-al do not stop to cook. The road does not wait.

\textbf{What does a child learn before they can walk?} To count the nine waves. Children are taught to watch the sea and count: one, two, three, four, five, six, seven, eight --- then stop. The ninth wave is the spirit's wave. It counts you. If you count it, you owe it.

\textbf{What is the first thing you notice about a stranger?} Their feet. A Lethai-al who has walked the circuit for many years has calloused soles cracked like dry earth. A stranger with soft feet has not walked far. A stranger with no feet is a ghost who lost the road.

\textbf{What is the most common crime?} Charm-theft --- stealing a tide-charm from a sleeping traveler. The thief gains the dead ancestor's favor for one night. The ancestor wakes. The ancestor is angry. The punishment is to carry the victim's burden for the rest of the circuit.

\textbf{What does no one talk about but everyone knows?} The drowned kingdom did not sink by accident. The forbidden name was spoken by a Lethai-al --- a woman who lost her child to the sea. She wanted the sea to remember. It remembered too well. Her name is not spoken. Her tide-charm hangs in the Sunken Gate, still wet.

\textbf{Where do people go when they die?} Into the tide-charms. The dead are not buried. Their names are carved onto shells and placed in oiled leather pouches. The living carry them. The dead walk with the living. When the pouch is lost, the dead become the Hollow.

\textbf{How do you apologize for a serious wrong?} You walk a crossroads circuit --- nine times around a crossroads shrine, counting your steps in eights. On the ninth circuit, you must speak the wrong aloud. If the wind carries your voice away, you are forgiven. If the wind throws it back, you are not. If the wind is silent, the Hollow is listening.

\textbf{What is the most important festival?} The Procession of the Ninth Wave (autumn). All Lethai-al bands walk to the shore at low tide. When the ninth wave crashes, they wade into the foam and whisper the names of those who drowned that year. The sea answers with a single wave. The wave is always the same height. The height is the number of names.

\textbf{What is the secret that could destroy a powerful person?} The Ninth Wave Herald is not a child. She is the first drowned --- the woman who spoke the forbidden name. She did not die. She became the ninth wave. She returns every year to point at the horizon. She is pointing at the day the sea will rise again. No one knows which day.
\end{almanac}

\newpage
\input{../generators/lethai_al_wanderers.tex}
\newpage

\chapter{Lethai-ar of Ayokha}
\label{chap:lethai-ar-ayokha}

\emph{The River-Bound Oath-Keepers.}

\noindent
{\Large
\textit{A witness cord is not a leash. It is a mirror. If you speak truly, you see yourself. If you lie, you see what you could have been. I have cut two cords in my life. I still dream of the faces in the silk. The river remembers every word spoken on its banks. We are merely the scribes.}

\vspace{0.5em}

My lamp has burned for nine years without being refilled. The flame is green. It flickers when a merchant lies about the weight of his silk. The river does not need oil. It needs witnesses. I sit on this wharf and watch. That is my offering.

\vspace{0.5em}

The Lethai-ar of Ayokha are the living bridges between mortals and the river spirits that launder the world's promises. Their ancestors were river-folk that ascended to the realm of the fair folk, and they now walk the middle path, a step between bodies of water and the world of men. They keep lanterns that never go out, record every contract spoken on the monsoon wharves, and bind their vows with silken cords dyed in river mud. Unlike their northern kin who swore to distant deities, the Lethai-ar serve \textbf{Benzaiten} (the river of words) or \textbf{Inari} (the fox's ledger). Their vows are physical cords, braided from their own hair and dyed with river mud. When a vow is fulfilled, the cord is burned. When it is broken, the cord tightens around the oath-breaker's wrist until blood is drawn. The Ayokhan taboo against the ninth cup is enforced by the Lethai-ar: the ninth cup is for the river spirits, and pouring it is an invitation to the Hollow.

\vspace{0.5em}

Never pour the ninth cup of tea or sake in a Lethai-ar's presence. The ninth cup is for the witness. Pour it, and you have invited the river to judge you. The river is not merciful.

\vspace{1em}

\noindent\textit{--- Vigil Lady Suyin, who has cut nine witness cords --- the ninth was her own}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
\textbf{What you notice first:} The smell of river water and oiled silk. The way every Lethai-ar carries a small oil lamp, even at noon. The constant murmur of contracts being witnessed --- a merchant's promise, a captain's oath, a lover's vow.

\textbf{The one rule that kills newcomers:} Never pour the ninth cup of tea or sake in a Lethai-ar's presence. The ninth cup is for the witness. Pour it, and you have invited the river to judge you. The river is not merciful.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Witness Cord}

Power here is measured in lanterns kept lit, cords tied, and the number of oaths witnessed. The Lethai-ar have no army, no treasury, no territory. They have the river, and the river remembers everything. Every contract in Ayokha requires a Lethai-ar witness to be legally binding. The God-King's astrologers may set the calendar, but the Lethai-ar hold the ledger. Their neutrality is absolute: they will witness a vow between enemies as readily as between allies. The price of a witness cord is not silver --- it is a promise to keep the river clean, to light a lamp at dusk, to speak no word that cannot be unsaid.

\subsection*{Starting Location}

A monsoon wharf on the great river, where silk barges are being loaded. A Vigil in gray robes holds a lantern of green flame: ``The merchant and the captain have made a vow. The captain says the silk is eight bales. The merchant says it is nine. The witness cord is tied. It will tighten around the liar's wrist. I do not need to know who is lying. The cord will tell me. But you --- you are not bound. Watch. Learn. And do not pour the ninth cup.''

\subsection*{The Almanac of Ordinary Days (Lethai-ar of Ayokha)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with river fish --- the fish are caught in the sacred currents, and the bones are returned to the water as an offering. The porridge is salted with evaporated river brine. It tastes of memory.

\textbf{What does a child learn before they can walk?} To tie the eight witness knots. Each knot represents a clause of a fair contract. The ninth knot is never tied --- it is left for the river to judge. A child who ties the ninth knot is taken to the step-pyramid and made to watch the green flame until they can recite every oath their family has ever sworn.

\textbf{What is the first thing you notice about a stranger?} Their lamp. A Lethai-ar's lamp is a public record of their witness. A stranger whose lamp burns green has kept every oath. A stranger whose lamp flickers has witnessed a broken vow. A stranger whose lamp is dark has been cut from the Vigil. Do not speak to them. The river may be listening.

\textbf{What is the most common crime?} Cord-forging --- creating a false witness cord that mimics the dye of a respected Vigil. The forger is not executed. Their hands are bound with a cord that tightens each time they speak. If they never lie, the cord will eventually loosen. If they lie once, it will cut.

\textbf{What does no one talk about but everyone knows?} The First Cord was not woven by a Lethai-ar. It was found --- already knotted, already witnessing --- floating down the river on a broken barge. No one knows who tied it. No one knows what oath it witnessed. The cord is kept in a lead-lined box at the bottom of the Well of Broken Oaths. The box is warm. The cord is still waiting.

\textbf{Where do people go when they die?} To the green flame. The dead are cremated, and their ashes are mixed with lamp oil. The oil is poured into the Vigil lanterns. The dead become the light. The light witnesses. The light does not lie.

\textbf{How do you apologize for a serious wrong?} You offer a new cord --- woven from your own hair, dyed with your own blood. The cord is tied between you and the wronged party. If the cord tightens, you are not forgiven. If it loosens, you are. If it breaks, the river will decide your fate. The river is patient.

\textbf{What is the most important festival?} The Reckoning of the Cords (spring). All witness cords that have been cut or broken are taken from the Well of Broken Oaths and burned in a barge on the river. The smoke rises. The river spirits watch. If a cord does not burn, the oath it witnessed is still unpaid. The debtor has one more year. If the same cord fails to burn twice, the Hollow comes.

\textbf{What is the secret that could destroy a powerful person?} The God-King's own witness cord is a forgery. He swore a vow to protect the Lethai-ar nine years ago. His cord was cut the next day. The Vigil knows. They have not told anyone because the cord that cut his was their own. The ninth cord that Lady Suyin cut was the king``s. She has been waiting for him to break his oath ever since. He will. They always do.
\end{almanac}

\newpage
\input{../generators/lethai_ar_river.tex}
\newpage

\chapter{Lethai-thora of Sihai}
\label{chap:lethai-thora}

\emph{The Silent Ledgers.}

\noindent
{\Large
\textit{We do not burn the dangerous scrolls. We bury them in plain sight, on shelves where no one thinks to look. The Emperor's Censor once stood in our archive and asked for a tax ledger. We gave him one. He did not notice the shelf behind him held the true history of his own grandfather's death. Silence is not emptiness. It is a door that opens only inward.}

\vspace{0.5em}

My grandmother tied this scroll with a silk cord. The cord has nine knots. The ninth knot is a secret she never told me. When I die, I will tie a tenth. Someone will untie it in a century. They will learn why the emperor's concubine wept on the night of the lantern festival. Some secrets are too heavy for one lifetime.

\vspace{0.5em}

The Lethai-thora were once courtiers of a celestial fey realm that spanned the Valewood and the eastern hills. When the old fey courts fell --- when the Alder King withdrew and the Hazel Queen sealed her gates --- many Lethai-thora fled east. They crossed the mountains with nothing but armfuls of scrolls and the memory of a civilisation that had forgotten how to die.

In the fertile river valleys of what would become Sihai, they found scattered agrarian tribes fighting over water and silence. The Lethai-thora did not conquer. They taught. They taught the tribes to write contracts instead of drawing blood, to measure grain instead of stealing it, to record debts instead of nursing grudges. Over generations, the tribes became a kingdom, the kingdom became an empire, and the empire became Sihai. The Lethai-thora became its memory.

Now they dwell in hidden pavilions, preserving ''forgotten ledgers`` that record what the Mandate of Heaven prefers to forget. Their knowledge is a burden: each secret they preserve slowly erases a piece of themselves. A Lethai-thora who guards a forbidden truth may find that she can no longer recall the color of her mother's dress --- a small price, she says, for keeping the ledger whole. They are no longer fey. They are no longer fully human. They are something in between: scribes of a story that has no ending, archivists of a past that never quite happened.

\vspace{0.5em}

Never touch a scroll without announcing your lineage. Even a false lineage is better than silence. The scrolls remember who opened them, and they will whisper your name to the Hollow.

\vspace{1em}

\noindent\textit{--- Keeper Lian, who has not spoken aloud in twenty years --- she writes on water}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
\textbf{What you notice first:} The smell of old parchment and incense. The way scrolls are stacked with their spines facing the wall --- so the words cannot see who reads them. The silence when a secret is spoken; even the insects stop chirping.

\textbf{The one rule that kills newcomers:} Never touch a scroll without announcing your lineage. Even a false lineage is better than silence. The scrolls remember who opened them, and they will whisper your name to the Hollow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Forgotten}

Power here is measured in scrolls sealed, secrets kept, and the length of a silence. The Lethai-thora have no king, no army, no treasury. They have archives, and the archives are their fortress. The Emperor's Censorate fears them because they cannot audit a secret that has no ink. The Mandate of Heaven is written in public; the Lethai-thora keep the unwritten ledger --- the one that records why dynasties fall. They are not rebels; they are witnesses. And a witness who never speaks is the most dangerous kind.

\subsection*{Starting Location}

A hidden pavilion behind a waterfall, where scrolls are stacked to the ceiling and the air is dry as bone. A Keeper of the Seventh Gutter looks up from a brass lamp: ''The Emperor's Quiet Chancellor has sent a messenger. He wants a scroll that does not exist. He wants the truth about his mother's death. The truth is written in water. It will vanish if we speak it. But you --- you are not bound by our silence. Will you carry a memory for us?``

\subsection*{The Almanac of Ordinary Days (Lethai-thora)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with dried fish --- the fish are smoked over fires fed with old scrolls. The smoke smells of forgotten contracts. The porridge is always lukewarm. The archives are never truly warm.

\textbf{What does a child learn before they can walk?} To read without moving their lips. A Lethai-thora child who whispers a secret aloud is sent to sit in the Chamber of the Ninth Scroll until they learn that silence is a shield. Some children never come back. They become scribes of the empty shelves.

\textbf{What is the first thing you notice about a stranger?} Their ink-stained fingers. A Lethai-thora has ink under every fingernail. A stranger whose fingers are clean has never copied a forbidden text. A stranger whose fingers are stained black has copied too many. A stranger whose fingers are stained red has copied a death warrant --- their own.

\textbf{What is the most common crime?} Context key theft --- stealing the carved stone that unlocks a scroll's meaning. The thief is not punished by law. They are given a blank scroll and told to write their own confession. The scroll will fill itself if the thief lies. The ink is their own blood.

\textbf{What does no one talk about but everyone knows?} The Ninth Scroll is not a scroll. It is a memory so dangerous that nine different keepers each sealed a fragment of it in their own minds. When the ninth keeper dies, the memory will be lost forever --- or will escape.

\textbf{Where do people go when they die?} To the Well of Forgotten Names. The dead are not buried. Their bodies are placed in the well, and the well swallows them. The water turns black. The black is the ink of their unwritten secrets. The well is bottomless. It is never full.

\textbf{How do you apologize for a serious wrong?} You offer a silence --- a period of total stillness, no breath, no heartbeat, no word. The silence is timed by a water timer. If you hold the silence for nine drops, you are forgiven. If you break it, you must speak the wrong aloud in the Chamber of the Ninth Scroll. The scroll will remember. The scroll does not forgive.

\textbf{What is the most important festival?} The Unrolling of the Ledgers (winter). Once a year, the keepers unroll a single forbidden scroll and read it aloud to each other --- but they read it in a whisper, and only the walls hear. The walls remember. The walls do not judge. The festival ends with all nine keepers drinking ink from a silver cup. The ink is their own memory. The next year, they will have forgotten.

\textbf{What is the secret that could destroy a powerful person?} The emperor's true lineage is recorded on a scroll that no living keeper has opened. The scroll is sealed with a cord tied by the first keeper. The cord has nine knots. The ninth knot is a lie --- the emperor's grandfather was not of the blood. The empire was founded by a concubine's son. The scroll is warm. It is kept in the First Pavilion. The First Pavilion is invisible. Only those who have forgotten nine names can see it.
\end{almanac}

\newpage
\input{../generators/lethai_thora_east.tex}
\newpage