\part{The Eastern Dolmis \& The Valewood}

% After \part{The Eastern Dolmis & The Valewood}
\begin{quote}
  \small
  \textit{``Every threshold has a keeper. Some keepers ask for salt. Some ask for silence. The Green Host asks for a name -- and it does not mean yours.''}
  \noindent --- Dusana of the Raven Road
  \end{quote}
% ------------------------------
% CHAPTER: THEONA
% ------------------------------
\chapter{Theona}
\emph{Three Greens, No Ninth.}

\noindent
{\Large
\textit{The wells here hold more than water---they hold the dreams and prayers of generations. To govern is to tend these sacred spaces where the living commune with what came before.}

\vspace{0.5em}

I am a Peat-Cutter. My barrow has a false bottom. I hide debts there---and bodies.

\vspace{0.5em}

The isles of Theona rise green and jagged from the Dolmis, their basalt cliffs wrapped in mists and moorland song. Nominally a duchy sworn to Viterra, in truth the Three Greens heed no crown but their own moots, no law but their taboos. The greatest of these is the silence of Nine: no bell tolls a ninth, no bridge counts that span, no oath dares its syllable.

\vspace{0.5em}

The Viterran sheriffs call it superstition. The Green Host calls it trespass. The difference is that the sheriffs will hang you. The Green Host will take your name, and you will wander the mist until you forget you ever had one.

\vspace{0.5em}

Never pour the ninth cup at a feast. Never speak the ninth name in a bargain. Never step where a step is missing. The Green Host rides for those who complete what should remain incomplete.

\vspace{1em}

\noindent\textit{--- Celia, Hedge-Witch Midwife, who has tied nine thousand knots and untied nine}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Oatcakes and heather ale---the oats are grown on terraced hillsides, the heather is harvested from the high moors. The ale is brewed in secret, by women who have forgotten more recipes than the men will ever learn. The men do not ask. The men are wise.

\textbf{What does a child learn before they can walk?} To count to eight. Theona children are taught numbers on their fingers---eight fingers, two thumbs. The thumbs are not counted. The thumbs are for pointing. Pointing is for warnings. Warnings are for the ninth thing.

\textbf{What is the first thing you notice about a stranger?} Their shadow. Theona is a land of mists and twilight. A stranger whose shadow is too short has been touched by the Host. A stranger whose shadow is too long is running from them. A stranger with no shadow is already taken. Do not speak to them. Do not meet their eyes. Do not offer them salt.

\textbf{What is the most common crime?} Breaking the ninth taboo---but breaking the taboo is not a crime, it is an invitation. The Host rides for those who pour the ninth cup, speak the ninth name, or take the ninth step. The Host does not arrest. The Host takes. The taken are never found. Their empty cups remain on the table. The cups are warm for three days.

\textbf{What does no one talk about but everyone knows?} The Green Host are not ghosts. They are the living---Theonans who broke the taboo and were taken. They ride with the King in Green, who was once a chieftain of the northern isle. He broke the ninth vow. He was not taken. He was crowned. His crown is made of thorns. His throne is a barrow. His court is the mist.

\textbf{Where do people go when they die?} To the barrows. Theona's dead are buried in long, low mounds that line the eastern cliffs. The mounds are planted with gorse and broom. The gorse blooms yellow in spring. The yellow is for mourning. The broom is for memory. The dead are not mourned. The dead are remembered. Remembering is enough.

\textbf{How do you apologize for a serious wrong?} You walk the Ninth Mile---a stretch of road that circles the central isle. The road is nine miles long. You must walk it at night, alone, with no light. If you reach the end by dawn, you are forgiven. If you do not, the Host will find you. The Host is always looking. The Host is always waiting.

\textbf{What is the most important festival?} The Feast of the Eighth Cup (autumn equinox). Every household pours eight cups of mead---one for each of the living generations. The ninth cup is poured onto the ground. The ground drinks. The ancestors listen. The ancestors do not answer.

\textbf{What is the secret that could destroy a powerful person?} The Three-Isles King is a changeling. The real king died in 798 AR, drowned in a peat bog while hunting. The changeling was placed by the Lady Beneath as part of a bargain. The bargain was for peace between the isles. The peace has held. The changeling has aged. The changeling's reflection does not.
\end{almanac}

\newpage
\input{../../generators/theona.tex}
\newpage

% ------------------------------
% CHAPTER: ZAKOV
% ------------------------------
\chapter{Zakov}
\emph{Salt \& Serpent.}

\noindent
{\Large
\textit{The tide here keeps better ledgers than any scribe, and the reefs remember every secret whispered above their depths. Power flows like the currents---those who understand them rule the waves.}

\vspace{0.5em}

I am a Dock-Rat. My tattoo is of the Sunken Quarter at low tide. The buildings are upside down.

\vspace{0.5em}

The pirate-city of the Dolmis coast isn't just a den of thieves---it's a corporation with a flag, and oh, how the Salt Prince loves playing king while the Syndicate Silent pulls the strings. Here, every manifest is a lie, every harbor a doorstep to a dozen conspiracies, and every fog-light a lure or a warning. The city breathes with the sea: at low tide, secrets exposed; at flood, the drowned walk the Sunken Quarter and knock on doors that should be locked.

\vspace{0.5em}

The Salt Prince's crown floats. The Syndicate's ledger is written in disappearing ink. The Kraken's Tongue lives in a bell tower that only exists at high tide. None of this is metaphor. All of it is true. All of it wants something from you.

\vspace{0.5em}

Never take a coin that tastes of brine. The Serpent's Mark is not a blessing. It is a debt that drowns.

\vspace{1em}

\noindent\textit{--- Ravel, Dock-Rat, who has never slept above the high-water mark}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salted fish and hard cheese---the fish is dried on the docks, the cheese is smuggled from the mainland. The cheese is always stale. The fish is always fresh. The contradiction is never explained.

\textbf{What does a child learn before they can walk?} To read the tide tables. Zakov children are taught the schedule of the sea before they learn their letters. A child who cannot read the tide is a child who will be caught by the flood. The flood is not water. The flood is the Syndicate. The Syndicate collects debts at high tide.

\textbf{What is the first thing you notice about a stranger?} Their coin-pouch. Zakov runs on silver---good silver, bad silver, silver that has been under the sea. A stranger whose pouch clinks with fresh-minted coin is a merchant. A stranger whose pouch is full of salt-crusted bits is a pirate. A stranger whose pouch is empty is a liar. Liars do not last long in Zakov. The sea takes them.

\textbf{What is the most common crime?} Piracy---but piracy is not a crime, it is a profession. The Syndicate licenses privateers. The privateers pay a tithe. The tithe funds the harbor. The harbor protects the ships. The ships are protected from pirates. The pirates are the Syndicate. The cycle is beautiful.

\textbf{What does no one talk about but everyone knows?} The Salt Prince is not a person. The title is held by a ring---a band of black coral that has been passed from captain to captain for three centuries. The ring is cursed. The ring's wearer cannot sleep. The ring's wearer cannot die. The ring's wearer cannot leave Zakov. The current Prince has worn the ring for forty years. He has not slept in forty years. He has not smiled in forty years. He has not left the palace in forty years.

\textbf{Where do people go when they die?} To the Sunken Quarter---the drowned district that lies beneath the harbor. The dead are weighted with chains and dropped into the canals. The canals empty into the sea. The sea returns the bodies at low tide. The bodies are collected by the Syndicate. The Syndicate sells them to the Salt-Prince's collectors. The collectors use them for---no one asks. No one wants to know.

\textbf{How do you apologize for a serious wrong?} You pay the Salt Tithe---a silver coin for every year of the wronged party's life. The coin must be a salt coin---one that has been under the sea for a full tide cycle. The coin is thrown into the harbor. If it sinks, you are forgiven. If it floats, you are not. If it floats back to shore, you are cursed. The curse is the Serpent's Mark. The Mark is a debt that drowns.

\textbf{What is the most important festival?} The Rising of the Tides (summer solstice). The harbor fills with luminescent algae. The algae glows green. The green is called ''the Kraken's Eye.`` The Kraken is said to rise on this night. The Kraken has not risen in living memory. The Kraken is patient. The Kraken is waiting.

\textbf{What is the secret that could destroy a powerful person?} The Pirate Queen is a spy. She was planted by the Kahfagian Admiralty forty years ago. Her mission was to destabilize the Syndicate. She has become the Syndicate. The Admiralty does not know. The Admiralty still sends her orders. She burns them. The ashes are mixed with her tea.
\end{almanac}

\newpage
\input{../../generators/zakov.tex}
\newpage

% ------------------------------
% CHAPTER: AELINNEL (GNOMES)
% ------------------------------
\chapter{Aelinnel (Aevrossa)}
\emph{Stone, Bough, and Bright Things.}

\noindent
{\Large
\textit{The law here flows like water over stone---it finds its own path, and woe to those who try to dam it with paper and pretense. The earth itself will testify against the unjust.}

\vspace{0.5em}

I am a Charcoal-Burner. My kilns tell us more than your books ever could. When the smoke curls left, the tide's with you. When it splits in three, best stay by your own hearth till it burns straight again.

\vspace{0.5em}

The gnomish homeland isn't just a place where they count steps---it's a place where they worship the count, and oh, how the Aelinnel love a good loophole in their beloved numerology. They count everything: steps, bells, offerings---because to them, precision isn't just virtue; it's leverage with a supernatural clause.

\vspace{0.5em}

The fae courts of the Thorn Lady and the Green Knight watch every threshold. The Stone Prince keeps the old pacts indexed in ledgers that breathe. And the Green Gate opens at neap tide, expecting exact change in truths. The Aelinnel don't just live between worlds---they arbitrate them.

\vspace{0.5em}

Never bring naked iron past a hawthorn arch unless it is gilded or named. Copper is polite, silver is opinionated, cold iron is an accusation. The hawthorn will remember which you carried.

\vspace{1em}

\noindent\textit{--- Tallow, Charcoal-Burner, who has never lost a sheep to the Hollow but has lost three helpers who did not listen}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Honey-cakes and mushroom tea---the honey is from the hawthorn groves, the mushrooms are harvested from the root-caverns. The cakes are baked in copper moulds. The moulds are shaped like animals. The animals are not edible. The shapes are for luck.

\textbf{What does a child learn before they can walk?} To count in eights. Aelinnel children are taught to group everything in eights---steps, breaths, heartbeats. The ninth is not a number. The ninth is a gap. The gap is where the Hollow sits. The Hollow is not a monster. The Hollow is a creditor. The debt is never paid in full.

\textbf{What is the first thing you notice about a stranger?} Their tools. Aelinnel carry copper---spoons, chisels, measuring rods. A stranger whose tools are iron is a barbarian. A stranger whose tools are silver is a priest. A stranger whose tools are gold is a fool. Gold does not ring true. Gold is for bribes. Bribes are for secrets.

\textbf{What is the most common crime?} Miscounting---but miscounting is not a crime, it is a geometric error. The error is corrected by the Stone Prince's ledgers. The ledgers are kept in the Hall of Sums. The Hall is guarded by clockwork sentinels. The sentinels do not sleep. The sentinels do not forgive. The sentinels count every breath.

\textbf{What does no one talk about but everyone knows?} The Stone Prince is not a prince. The title is held by a committee of six elders who have never been seen together. They communicate through copper messages left in the Hall of Sums. The messages are written in a code that changes with the tides. The code has not been broken in three centuries. The code is not meant to be broken. The code is meant to be trusted.

\textbf{Where do people go when they die?} To the tide-gates---stone arches at the edge of the sea. The dead are laid on the threshold. The tide comes in. The tide takes them. The gates are older than the Aelinnel. The gates were built by something else. The gates are still watching.

\textbf{How do you apologize for a serious wrong?} You offer a counting---a public recitation of your wrong in the Stone Prince's court. The court is a circle of copper chairs. The chairs are cold. The wrongdoer stands in the center. The elders count the syllables of your apology. If the count is even, you are forgiven. If the count is odd, you must leave an offering. The offering is measured in copper. The copper is weighed against your breath. The scales are never balanced.

\textbf{What is the most important festival?} The Opening of the Green Gate (autumn equinox). The gate appears at low tide, a stone arch that leads to---no one knows. The Aelinnel do not enter. They leave offerings: copper, honey, a single true name. The gate accepts. The gate closes. The gate waits.

\textbf{What is the secret that could destroy a powerful person?} The Green Knight is not a knight. She is the gatekeeper---the last survivor of the people who built the tide-gates. She wears a helm of polished brass. Her face has not been seen in living memory. Her voice is a recording. Her sword is a key. The key opens the gate. The gate does not want to be opened.
\end{almanac}

\newpage
\input{../../generators/aelinnel.tex}
\newpage

% ------------------------------
% CHAPTER: AELAEREM (HALFLINGS)
% ------------------------------
\chapter{Aelaerem}
\emph{Hearth \& Hollow.}

\noindent
{\Large
\textit{A proper harvest feast requires three things: the finest cider aged in the right barrel, a guest-loaf baked with intention, and the wisdom to know when to invite the quiet ones to sit by the fire.}

\vspace{0.5em}

I am a Miller. My geese know a stranger's step from a mile off, and they've never been wrong. One is named ''Stranger.`` It has never been wrong.

\vspace{0.5em}

The halfling downs aren't just a place where they keep the Hollow at bay---they're a marketplace for fear, and oh, how the Aelaerem love a good bargain on safety. They use red threads and offerings not out of piety, but because they know the Hollow's attention is a commodity---and they'll pay dearly to avoid it.

\vspace{0.5em}

A red door means bread, salt, and one safe night if you come honest. A cup-mark stone on the verge accepts butter as a prayer. The watch-geese know your step before you do, and the scarecrow turns to face the road when the Hollow is hungry. The Neighbors---the things under the barrows---are not enemies. They are neighbors. And neighbors, if treated politely, will stay on their side of the hedge.

\vspace{0.5em}

Never refuse an offered cup. The Hollow counts refusals as invitations. Pour the ninth cup for the empty chair, and you may feed it---or feed it.

\vspace{1em}

\noindent\textit{--- Bram, Miller of the Watch-Geese, who has never been surprised}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Porridge with honey and a pinch of salt---the salt is for the Hollow. Every Aelaerem household leaves a pinch on the threshold each morning. The Hollow does not eat. The Hollow counts. The salt is an offering. The offering is a reminder.

\textbf{What does a child learn before they can walk?} To tie a red thread. Aelaerem children are given a length of red wool at birth. By age three, they can tie the eight hearth-knots: welcome, farewell, protection, memory, warning, promise, apology, and the ninth---the one that is never tied. The ninth knot is for the Hollow. The Hollow does not tie knots. The Hollow waits for you to tie them yourself.

\textbf{What is the first thing you notice about a stranger?} Their shadow. Aelaerem shadows are long and thin, even at noon. A stranger whose shadow is short has been touched by the Neighbors. A stranger whose shadow is wrong---too many limbs, too few---is a Neighbor. Do not invite them in. Do not refuse them bread. The law is clear: feed them, but do not ask their name.

\textbf{What is the most common crime?} Offering the ninth cup---but the ninth cup is not a crime, it is a negotiation. The pourer offers a drink to the empty chair. The Hollow may accept. If it does, the house is blessed for a year. If it does not, the house is marked. The mark is a red thread tied to the door. The thread is never removed. The house is never sold.

\textbf{What does no one talk about but everyone knows?} The Apple-Matron is a title, not a person. The Matron is the oldest woman in the village. When she dies, the second-oldest takes her place. The Matron's memory is the village's memory. The Matron forgets nothing. The Matron's memory is fading. The village is forgetting. The forgetting is the Hollow's doing.

\textbf{Where do people go when they die?} To the barrows---the low, green mounds that dot the downs. The dead are buried in family plots, marked with cup-stones. The cup-stones are filled with rainwater. The rainwater is changed weekly. The water is for the dead. The dead are not thirsty. The dead are counting.

\textbf{How do you apologize for a serious wrong?} You leave a red thread on the wronged party's doorstep. The thread must be tied in a specific knot---the knot of apology, which is also the knot of debt. The wronged party may accept the thread or cut it. If they accept, the debt is recorded. If they cut it, the debt is unpaid. The unpaid debt grows. The growing debt attracts the Hollow.

\textbf{What is the most important festival?} The Harvest Home (autumn). The last sheaf of grain is tied with red thread and left in the field. The Hollow takes it. The Hollow does not take anything else. The Hollow is patient. The Hollow will wait for the ninth cup.

\textbf{What is the secret that could destroy a powerful person?} The Pale Shepherd is not a shepherd. It is a title held by the Hollow itself. The Shepherd appears as a figure in gray robes. Its face is a mask of bone. Its voice is a recording of the last words of the person who saw it last. That person is always dead. The recording is always accurate.
\end{almanac}

\newpage
\input{../../generators/aelaerem.tex}
\newpage

% ------------------------------
% CHAPTER: THE VALEWOOD
% ------------------------------
\chapter{The Valewood}
\emph{Empire Under Leaves.}

\noindent
{\Large
\textit{The forest here remembers empires that crumbled to dust, and its laws are older than your kingdoms. To walk these paths is to enter into covenant with powers that measure time in seasons, not years.}

\vspace{0.5em}

I am a Pathweaver. My cord has eight knots. The ninth is the one she never ties.

\vspace{0.5em}

The Valewood isn't just ancient---it's a graveyard of empires, and oh, how the Lethai-ar ghosts love a good deal on forgotten magic. They don't just rule as echoes; they trade in them. The air hums with half-remembered songs, and the ruins bleed their mysteries into root and stream. Fae courts linger in clearings that shift with moonlight, beast-kin patrol moss-choked cairns, and echoes of imperial shades still march patrols on vanished streets.

\vspace{0.5em}

The Hazel Queen speaks only in conditional tense. The Alder King's crown is a noose of thorns. The Huntsman's face is a stag's skull, and his voice is the wind in the leaves. The forest does not forgive arrogance. It expects humility, patience, and the willingness to trade a piece of yourself for safe passage.

\vspace{0.5em}

Never eat the fruit of a tree that has a face. The forest does not forget its debtors, and the first bite is always the price of admission. The second bite is the interest.

\vspace{1em}

\noindent\textit{--- Pathweaver Sila, who knows every star-road but will not speak their names}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Acorn bread and wild honey---the acorns are leached in stream-water for three days, the honey is taken from hives in the high branches. The bees are sacred. The bees are not to be harmed. A beekeeper who harms a bee is cursed. The curse is the stinging. The stinging is endless.

\textbf{What does a child learn before they can walk?} To name the trees. Valewood children are taught the names of the eight sacred groves: Hazel, Alder, Oak, Rowan, Ash, Thorn, Yew, and the ninth---the one that has no name. The ninth grove is where the Hollow sleeps. The Hollow is not to be woken.

\textbf{What is the first thing you notice about a stranger?} Their way-cord. Every Valewood traveler carries a knotted cord---one knot for each mile they have walked. A stranger whose cord has nine knots has walked the forbidden path. A stranger whose cord is knotted with red thread is a Pathweaver. A stranger with no cord has no way. The forest will find them. The forest will lead them in circles until they forget why they came.

\textbf{What is the most common crime?} Breaking the star-road---but the star-roads are not roads, they are agreements. A traveler who strays from the path is a traveler who breaks the covenant. The covenant is older than the Valewood. The covenant was made between the first trees and the first stars. The stars do not forgive. The trees do not forget.

\textbf{What does no one talk about but everyone knows?} The Hazel Queen is not the queen. She is the voice of the forest---a title that passes from dryad to dryad with each turning of the season. The current Queen has held the title for three centuries. She is not a dryad. She is a ghost---the last survivor of the empire that fell beneath the leaves. She remembers the cities. She remembers the stones. She remembers the names of the dead. The dead do not remember her.

\textbf{Where do people go when they die?} To the star-roads---the paths that lead between the trees and between the worlds. The dead are not buried. They are laid on stone altars at the edges of the groves. The stars pull them upward. The trees pull them downward. The dead choose. The dead always choose the trees. The trees remember.

\textbf{How do you apologize for a serious wrong?} You offer a name---your own or another's. The forest accepts the name. The forest weaves it into the bark of the oldest oak. The oak is called the Witness. The Witness has stood for nine thousand years. The Witness has never forgotten a single name. The Witness will not forget yours.

\textbf{What is the most important festival?} The Turning of the Leaves (autumn). The trees shed their colors in a single day. The leaves carpet the forest floor. The leaves are gathered and burned in the high clearings. The smoke carries the year's debts to the stars. The stars answer with light. The light is forgiveness. The light is also a warning.

\textbf{What is the secret that could destroy a powerful person?} The Alder King is not a king. He is the forest's warden---a title passed from fallen empire to fallen empire. The current King is a Lethai-ar who was born in the year the last empire fell. He has not aged since. He does not eat. He does not drink. He does not sleep. He walks the star-roads at night. The star-roads are empty. The dead have stopped walking.
\end{almanac}

\newpage
\input{../../generators/valewood.tex}
\newpage