\part{The Northern Reaches}
% After \part{The Northern Reaches}
\begin{quote}
\small
\textit{``The mountain counts your breaths. The frost writes your name in ice. In the north, the dead do not wait -- they walk, and they remember every promise you made to keep warm.''}
\noindent --- Dusana of the Raven Road
\end{quote}
% ------------------------------
% CHAPTER: AELER
% ------------------------------
\chapter{Aeler}
\emph{Crowns \& Under-Vaults.}

\noindent
{\Large
\textit{The mountain remembers every word carved in its stone, every seal pressed into its walls. To govern here is not to command, but to harmonize with the ancient songs that still echo in the deep places.}

\vspace{0.5em}

I am an Under-Mason. I can hear stone shift a room away. When the rock starts singing a different tune, you best listen. Three knocks in the morning means all's well, but three knocks at night means something's coming through the walls that wasn't meant to walk.

\vspace{0.5em}

My homeland. The place where we don't trust the surface, but we do trust the weight of a well-placed keystone---and oh, how the surface loves to test that trust. Our breath-bonds aren't just currency---they're the hardest thing in the peaks, forged in the sighs of miners who knew too much. We sell them to Vilikari jarls who pay in furs stained with brother's blood, to Ykrul khagans who weigh the value in warhorses and broken oaths. Once, we sold one to the Veiled Sisters. They paid in secrets---which is another way of saying they paid in future leverage.

\vspace{0.5em}

Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one---a breath, a memory, a shadow. My grandfather reached nine once. He lived another thirty years, but he never cast a shadow again.

\vspace{1em}

\noindent\textit{--- Thromm Stone-Tongue, Under-Mason, who has never counted to nine}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Stone-bread and cave cheese---the bread is baked on heated stone slabs, the cheese is aged in sealed galleries where the temperature never varies. The cheese has been aging for three generations. It is an heirloom. It is also barely edible.

\textbf{What does a child learn before they can walk?} To count their breaths. Aeler children wear breath-counters---small brass bells on leather cords. Eight breaths, ring the bell. Nine breaths, run. A child who cannot count is a child who will not survive the winter.

\textbf{What is the first thing you notice about a stranger?} Their bell. Every Aeler carries a bell tuned to their breath. A stranger's bell will ring at a different pitch---the mountain's lungs are calibrated to local frequencies. A mismatched bell is a warning. A silent bell is a threat.

\textbf{What is the most common crime?} Air theft. In the deep vaults, ventilation is shared. A sealed door, a blocked vent, a diverted shaft---these are thefts of breath, punishable by breath-debt. The guilty party must donate their own air ration to the victim. The donation is measured in days. The days are counted by bells.

\textbf{What does no one talk about but everyone knows?} The True Masons are not neutral. They are the Vault-Queen's eyes. Every ''independent`` repair, every ''neutral`` inspection, every ''routine`` audit---the True Masons are watching. They do not report to the Queen. They are the Queen. They have always been the Queen.

\textbf{Where do people go when they die?} To the Deep Vaults---the oldest galleries, where the air is thin and the stone is cold. The dead are not buried. They are sealed in niches carved into the walls, their names chiselled above their heads. The niches are reused every century. The names are never erased. The dead share walls with strangers. The strangers are patient.

\textbf{How do you apologize for a serious wrong?} You offer a keystone---a load-bearing stone from your own holding. The stone is installed in the wronged party's wall. If it holds, you are forgiven. If it cracks, the debt stands. If it crumbles, your holding collapses. Apology is architecture. Architecture is law.

\textbf{What is the most important festival?} The Ringing of the Underways (deep winter). Every bell in every hold is rung at the same hour---eight times, no more. The sound travels through the stone, a heartbeat that reminds the mountain that the living are still here. The dead listen. The dead do not ring back.

\textbf{What is the secret that could destroy a powerful person?} The Vault-Queen is a construct. She was assembled in 723 AR, after the real queen died in a cave-in. Her heart is a keystone. Her breath is a bellows. Her voice is a recording. The True Masons maintain her. They have the schematics. The schematics are missing three pages. The missing pages describe how to turn her off.
\end{almanac}

\newpage
\input{../generators/aeler.tex}
\newpage

% ------------------------------
% CHAPTER: THE MISTLANDS
% ------------------------------
\chapter{The Mistlands}
\emph{Bells, Salt, and Breath.}

\noindent
{\Large
\textit{The bells here do more than mark time---they weave the very fabric of reality, keeping the mist at bay and the dead in their proper places. Each chime is a thread in the pattern that holds the world together.}

\vspace{0.5em}

I am a Bell-Warden. I refuse to write down the ninth note. It lives only in my memory---and that memory is fading.

\vspace{0.5em}

The Mistlands are a protectorate of the Aeler in name only. The fog here doesn't just hide the land; it hides the law, and oh, how the wraiths love that ambiguity. The bell-lines aren't just defenses---they're the only thing keeping the wraiths from dragging whole villages into the mist. The Aeler built them. The Aeler charge for them. The Aeler will leave when the money runs out. And when that happens, the fog will remember every single soul they abandoned.

\vspace{0.5em}

Never answer a voice that calls your name from the fog. The mist collects names, and once it has yours, the ferryman will always know where to find you. I've seen men answer. I've watched them walk into the white, still talking, still smiling, until their voices came back from a direction they hadn't gone.

\vspace{1em}

\noindent\textit{--- Warden Noah, who has seen the ninth bell rung once and will not speak of it}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salt fish and sour bread---the fish is smoked over peat fires, the bread is leavened with fog-damp yeast. The yeast is caught in cloth traps hung from the bell-towers. The traps are emptied at dawn. The yeast is not alive. It is not dead either.

\textbf{What does a child learn before they can walk?} To ring a bell. Every Mistlands child is given a small bronze bell at birth. By age three, they can ring the eight notes of the ward-chime in their sleep. The ninth note is never taught. The ninth note is discovered. Discovery is fatal.

\textbf{What is the first thing you notice about a stranger?} Their salt-pouch. Every Mistlander carries a leather pouch of consecrated salt. A stranger without a pouch is either a fool or a wraith. A stranger whose pouch is empty is a fool who is about to become a wraith.

\textbf{What is the most common crime?} Ward-salt theft. The Aeler control the salt pans, and the salt is the only thing that keeps the dead from rising. A stolen pouch can buy a week's protection. The price is a year's debt to the Aeler. The debt is collected in labor. The labor is collecting salt. The salt is stolen. The cycle is perpetual.

\textbf{What does no one talk about but everyone knows?} The Bell-Wardens are not chosen---they are volunteered. A village that loses a warden must send a replacement to the Aeler training holds. The holds are in the mountains. The mountains are cold. The trainees do not always return. The ones who do return do not speak of what they learned. Their bells ring at a different pitch.

\textbf{Where do people go when they die?} To the fog. Bodies are not buried in the Mistlands---the ground rejects them. Corpses are carried to the edge of the bell-line and left for the mist. The mist takes them. Sometimes, at dusk, you can see them walking the lanes. They do not speak. They do not need to. Their presence is a question.

\textbf{How do you apologize for a serious wrong?} You ring the Bell of Atonement---a cracked bronze bell that hangs in every village square. The bell rings once for each wrong confessed. If the bell rings true, you are forgiven. If it cracks further, you are not. If it shatters, you are banished to the fog. The banished never return. Their voices still answer from the white.

\textbf{What is the most important festival?} The Rehanging of the Lines (autumn). Every bell-line is taken down, cleaned, and rehung. The work takes three days. The villages are silent for three nights. The dead listen. The dead wait. The dead do not ring.

\textbf{What is the secret that could destroy a powerful person?} The Legate of the Mists is a wraith. She died in 803 AR, during a bell-line failure. The Aeler rebuilt her from fog and memory. She does not know she is dead. She attends council meetings. She signs documents. Her signature fades after sunset. The documents are refiled every morning.
\end{almanac}

\newpage
\input{../generators/mistlands.tex}
\newpage

% ------------------------------
% CHAPTER: YKRUL
% ------------------------------
\chapter{Ykrul}
\emph{Wolf Standards, Winter Camps.}

\noindent
{\Large
\textit{The steppe remembers every oath sworn beneath its endless sky, and the wolf-standards carry the honor of clans that have outlasted empires. To lead here is to keep the old ways sharp as winter steel.}

\vspace{0.5em}

I am a Herd-Scout. I can read a trail by the way the grass bends---and a lie by the way a rider sits. The Ykrul don't measure wealth in silver; we measure it in hoofbeats, in the snort of a warhorse at dawn, in the weight of a promise kept through winter.

\vspace{0.5em}

We do not build monuments. We build relationships, each one a stone in an invisible cairn that stretches from the Dolmis to the Aberderrin. A camp is a weave of favors owed and honor satisfied. A feud is a slow equation: blood in, blood out, with interest compounded by winter.

\vspace{0.5em}

The Khagan's Court doesn't rule from a throne---it rules from a saddle. Our ''laws`` are the Thirty-Six Sayings of the Road, and the thirty-seventh is the one you learn alone, on a night when the stars blink out and the wolf does not howl.

\vspace{0.5em}

Never ride between two wolves at dusk. The space between them is a blind angle---a route with no exit, a debt with no payment. The wolves know this. They choose the gap deliberately.

\vspace{1em}

\noindent\textit{--- Sogata, Herd-Scout, who has never been lost but has been found many times}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Fermented mare's milk and dried meat---the milk is drunk warm from the skin, the meat is sliced thin and dried on saddle-pommels. The meat carries the taste of horse sweat. The milk carries the taste of the steppe. Both are acquired. Both are necessary.

\textbf{What does a child learn before they can walk?} To sit a horse. Ykrul children are placed on a horse before they can stand. The horse is old, patient, and knows the child's weight. A child who falls is a child who learns to hold tighter. A child who cannot ride is a child who cannot eat.

\textbf{What is the first thing you notice about a stranger?} Their braids. Ykrul warriors braid their hair with clan colors---red for the Wolf Clan, black for the Iron Horse, gray for the Salt Steppe. A stranger whose braids are cut has been shamed. A stranger whose braids are dyed black has killed a clan-mate. A stranger with no braids has no clan. A stranger with no clan is prey.

\textbf{What is the most common crime?} Horse theft. A stolen horse is a stolen relative. The price is blood---the thief's or the owner's. The courts are the council fires. The evidence is the horse's memory. Horses remember faces. Horses remember smells. Horses do not lie.

\textbf{What does no one talk about but everyone knows?} The Khagan is not chosen by blood. He is chosen by the horse. Every five years, the clan chiefs lead their best stallions to the sacred ring. The horse that steps forward first chooses the next Khagan. The horse has not stepped forward in twenty years. The council chooses by silence. The silence is growing loud.

\textbf{Where do people go when they die?} To the sky. Ykrul dead are placed on open platforms, wrapped in felt, and left for the wind and the birds. The bones are collected after a year and ground into powder. The powder is mixed with clay and fired into beads. The beads are worn by the dead's descendants. The beads are warm. The warmth is memory.

\textbf{How do you apologize for a serious wrong?} You offer a hostage string---a braided cord that ties your wrist to the wronged party's. The cord is cut when the apology is accepted. If the cord is cut before the apology is finished, you are not forgiven. If the cord is never cut, you are bound. The binding is lifelong. The binding is called ''friend.``

\textbf{What is the most important festival?} The First Grass (spring equinox). The herds are released onto the new pasture. The riders race to the horizon and back. The winner names the summer camps. The loser tends the sick animals. The sick animals are always the loser's. The loser always wins next year.

\textbf{What is the secret that could destroy a powerful person?} The Sky-Speaker is a fraud. She cannot call storms. She cannot read omens. She cannot speak to the dead. She is a horse-witch who learned to read the clouds. The clouds are reliable. The Khagan does not know. The Khagan does not want to know.
\end{almanac}

\newpage
\input{../generators/ykrul.tex}
\newpage

% ------------------------------
% CHAPTER: VILIKARI
% ------------------------------
\chapter{Vilikari}
\emph{Laurels \& Longhouses.}

\noindent
{\Large
\textit{We are the bridge between empire and wilderness, the marriage of law and custom. Our strength lies not in choosing one path but in walking both with equal surety.}

\vspace{0.5em}

I am a Hearth-Mother. I have never lost a quarrel. My voice is soft, my memory longer than the Foedus Stone.

\vspace{0.5em}

The Vilikari straddle the border like a drunkard on a fence---half empire, half steppe, all headache. Their foedus treaties aren't just agreements---they're a tragicomic mess, half carved in marble by emperors who forgot to pay the sculptors, half braided in string by chieftains who forgot which knot meant ''peace.`` And yet, somehow, the treaties hold. Not because of the marble. Not because of the string. Because both sides are too exhausted to start the argument over again.

\vspace{0.5em}

Never invoke only one law. The mallus (oak-ring) and the basilica (marble bench) must both be honoured. Choose one, and the other will find you---usually at midnight, usually with a writ, usually with a sword.

\vspace{1em}

\noindent\textit{--- Hearth-Mother Vesna, who has fed three kings and forgotten none of their debts}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Oatmeal with dried berries---the berries are harvested from the border scrub, where the empire's orchards meet the steppe's wild brambles. The berries taste of both. The oatmeal tastes of neither. The oatmeal is filler. The berries are the point.

\textbf{What does a child learn before they can walk?} To speak both languages---Imperial Utaran and the steppe cant. A child who speaks only one is a child who will never cross the border. A child who crosses the border is a child who will come back with stories. A child with stories is a child who will be listened to. Listening is power.

\textbf{What is the first thing you notice about a stranger?} Which side of the road they ride on. The empire's roads are measured from the center. The steppe's paths are measured from the edge. A stranger who rides in the center is an imperial. A stranger who rides at the edge is a nomad. A stranger who rides in the ditch is a spy. The ditch is not a road. The ditch is a message.

\textbf{What is the most common crime?} Treaty-splitting---selling the same rights to both sides. A Vilikari factor will promise river access to an imperial merchant and grazing rights to a Ykrul clan, then collect from both and disappear before the paperwork collides. The crime is not in the selling. The crime is in getting caught before you cross the border.

\textbf{What does no one talk about but everyone knows?} The Foedus Stone is a copy. The original was stolen in 712 AR by a Vilikari trader who owed debts on both sides. She melted it down and sold the marble dust as ''treaty-strengthening powder.`` The powder was mostly chalk. The treaties are still holding. The traders are still selling.

\textbf{Where do people go when they die?} To the border. The Vilikari bury their dead at the edge---close enough to the empire for the priests to pray, close enough to the steppe for the shamans to chant. The dead are planted face-up, so they can see both skies. The skies are different colors. The dead do not complain.

\textbf{How do you apologize for a serious wrong?} You invoke the mallus---the oak-ring court. The wronged party sits inside the ring. The wrongdoer circles it three times, speaking their apology. On the third circuit, the elders judge. If the apology is accepted, the wrongdoer is admitted into the ring. If it is rejected, the ring closes. The wrongdoer is outside. Outside is the lawless land. The lawless land is where the wolves are.

\textbf{What is the most important festival?} The Renewal of the Foedus (summer solstice). The stone is washed with milk and honey. The milk is from a Ykrul mare. The honey is from an imperial apiary. The mixture is drunk by the elders. The elders speak the treaty aloud. The treaty is long. The elders are old. The ceremony takes all day.

\textbf{What is the secret that could destroy a powerful person?} The Foedus Stone has a crack. The crack was caused by a hammer in 799 AR, during a dispute over horse tariffs. The stone was repaired with clay and prayer. The clay is drying. The prayer is fading. The crack is spreading. When the stone splits, the treaty splits with it. No one knows what happens then. No one wants to find out.
\end{almanac}

\newpage
\input{../generators/vilikari.tex}
\newpage

% ------------------------------
% CHAPTER: BLACK BANNERS
% ------------------------------
\chapter{Black Banners}
\emph{Condotta \& Crowns.}

\noindent
{\Large
\textit{A contract is only as good as the next payday, and a grudge is only as sharp as the sword backing it. In this game, we sell steel by the season and vendettas by the year.}

\vspace{0.5em}

I am a Veteran Sergeant. I've seen three captains rise and fall in the time it takes to grow a beard, but the mud stays the same and the pay comes when it pleases.

\vspace{0.5em}

Mercenary companies sell steel to the highest bidder---and oh, how they love a good spreadsheet. The Banner-Captains don't just want payment; they want proof they're getting their money's worth, which is why they pay in safe-conducts that are good until the next battle\dots and utterly worthless after, once the smoke clears and the ledgers need balancing.

\vspace{0.5em}

The camp sings when payday comes. The camp falls silent when the ledger is written in blood. I've heard both. I prefer the singing. At least the singing means you're still alive.

\vspace{0.5em}

Never break a Bell Dawn. Three bells at sunrise means bury the dead, trade prisoners, fix the wells, no ambushes. Violating it is how legends end badly---and how captains end hanging.

\vspace{1em}

\noindent\textit{--- Old Sorka, Veteran Sergeant, who counts scars like coins and spends them on whiskey and mercy}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Hardtack and camp stew---the stew is whatever was caught, killed, or traded for before dawn. The cook is called ''The Alchemist.`` No one knows their real name. No one asks.

\textbf{What does a child learn before they can walk?} To count coin. Camp children are taught to tally pay shares before they can read. A child who miscounts is a child who will go hungry. A child who goes hungry is a child who learns to count faster.

\textbf{What is the first thing you notice about a stranger?} Their colors. Every Black Banner company has its own banner---lantern, thorn, bell, fishbone, broken scale. A stranger whose colors are faded has been running. A stranger whose colors are too bright has never fought. A stranger with no colors has no company. A stranger with no company is a deserter. A deserter is a liability. A liability is a corpse.

\textbf{What is the most common crime?} Desertion---but desertion is not a crime, it is a negotiation. A soldier who wants to leave will sell their gear to the quartermaster, sell their secrets to the enemy, and sell their silence to the captain. The price is negotiable. The price is always ''one more season.``

\textbf{What does no one talk about but everyone knows?} The Black Colonel is not a person. The Colonel is a title held by a wax seal. The seal is kept in a lead-lined box. The box is carried by a mute lieutenant. The lieutenant signs documents with the Colonel's thumb-print. The thumb-print is a forgery. The forger is the lieutenant. The lieutenant has no thumbs.

\textbf{Where do people go when they die?} To the pyre. Camp dead are burned in a bonfire of broken spears and banner poles. The ashes are scattered by the quartermaster. The quartermaster charges a fee. The fee is deducted from the dead's final pay. The final pay is never enough.

\textbf{How do you apologize for a serious wrong?} You stand a watch. The wronged party chooses the hour---midnight, dawn, the dead of winter. You stand alone. If you survive, you are forgiven. If you do not, the debt is paid. The camp does not mourn. The camp does not forget.

\textbf{What is the most important festival?} Payday---but payday is not a day, it is a negotiation. The captain promises payment on the first of the month. The quartermaster promises payment when the supply train arrives. The soldiers promise loyalty until the payment comes. The payment never comes. The loyalty never ends.

\textbf{What is the secret that could destroy a powerful person?} The Banner-Captain's contract is written in disappearing ink. The ink was supplied by a Sidhi forger who was paid in promises. The promises were not kept. The forger is now a camp follower. The forger knows where the contract is hidden. The forger is waiting for the right offer.
\end{almanac}

\newpage
\input{../generators/black_banner.tex}
\newpage

% ------------------------------
% CHAPTER: THE LINNS
% ------------------------------
\chapter{The Linns}
\emph{Skerries \& Storm-Oaths.}

\noindent
{\Large
\textit{The sea keeps better counsel than any scribe, and the storm remembers every oath sworn upon its winds. To rule these shores is to dance between their moods and never miss a step.}

\vspace{0.5em}

I am a Net-Wife. I tie a knot for every storm I've survived. My belt has thirty.

\vspace{0.5em}

North of the Aberderrin Sea, the Linns don't just raid and trade---they do it with the fervour of the converted, swearing oaths to storm and sea that aren't just words---they're leverage with a supernatural clause. They pay in stories, not silver, because to them, a tale well-told is worth more than a chest of gold\dots especially when it's true.

\vspace{0.5em}

The jarls measure their wealth in ships. The skalds measure their wealth in verses. The dead measure their wealth in how many living remember their names. I measure mine in knots. Each knot is a promise I kept to the sea. The sea remembers. The sea does not forgive forgetting.

\vspace{0.5em}

Never name a drowned person aloud while underway. The sea collects names, and if it hears you speaking one it doesn't recognize, it may reach up to correct the ledger.

\vspace{1em}

\noindent\textit{--- Astrid, Net-Wife, who has never lost a sailor to the mist but has lost three to bad stories}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Seal blubber and flatbread---the blubber is chewy, the bread is dense, and both are stored in the same barrel. The barrel is shared with the crew. The crew shares everything. The crew is family. Family is obligation.

\textbf{What does a child learn before they can walk?} To tie a knot. Linn children are given a length of rope at birth. By age five, they can tie the eight sea-knots: clew, clove, figure-eight, reef, sheet, stopper, timber, and the ninth---the one that slips when pulled. The ninth knot is for emergencies. The ninth knot is for funerals. The ninth knot is for when the sea asks for a life.

\textbf{What is the first thing you notice about a stranger?} Their breath. Linn sailors chew kelp to sweeten their breath---the kelp is harvested from sacred beds, blessed by the Volva. A stranger whose breath smells of salt has been at sea. A stranger whose breath smells of nothing has not. A stranger whose breath smells of copper has spoken a false oath. The sea knows. The sea will answer.

\textbf{What is the most common crime?} Oath-breaking---but oath-breaking is not a crime, it is a storm. The oath-breaker's ship will founder. The oath-breaker's nets will tear. The oath-breaker's children will be born with webbed fingers. The punishment is not imposed by the jarl. The punishment is imposed by the sea. The sea is patient. The sea does not forget.

\textbf{What does no one talk about but everyone knows?} The Sea-Queen is not a queen. She is a figurehead---literally. Her body is carved from the keel of a ship that sank in 612 AR. Her voice is a recording. Her throne is a navigation table. The jarls pretend. The jarls have always pretended. The pretending keeps the peace.

\textbf{Where do people go when they die?} To the waves. Linn dead are placed in boats and set adrift, weighted with stones. The stones are carved with the dead's name. If the boat sinks before it reaches the horizon, the dead is accepted. If it reaches the horizon, the dead is rejected. If it returns to shore, the dead is cursed. The cursed wash up on beaches. The beaches are littered with their bones.

\textbf{How do you apologize for a serious wrong?} You offer a storm-oath. You stand on a cliff during a gale and swear your apology to the wind. If the wind carries your words, you are forgiven. If the wind throws you into the sea, you are not. The sea does not give second chances.

\textbf{What is the most important festival?} The Launching of the Longships (spring thaw). The ships are dragged from their winter shelters and blessed with blood and mead. The blood is from a sacrificed goat. The mead is from the jarl's own stores. The ships are named after ancestors. The ancestors are listening. The ancestors approve.

\textbf{What is the secret that could destroy a powerful person?} The Volva of the Mist is a fraud. She cannot see the future. She reads the weather---the pressure, the wind, the taste of salt on the air. The sailors know. The sailors do not care. A weather-reader is more useful than a seer. The Volva knows this. The Volva uses this. The Volva is never wrong.
\end{almanac}

\newpage
\input{../generators/linns.tex}
\newpage