\part{The Spaces Between}
% After \part{The Spaces Between}
\begin{quote}
  \small
  \textit{``The Ways Between do not measure miles. They measure choices. A wrong turn might lead to yesterday. A right turn might lead to a version of yourself who made a different decision -- and is still alive to regret it. The mist does not hide landmarks. It hides destinations.''}
  \noindent --- Dusana of the Raven Road, at a crossroads that did not exist on any map
  \end{quote}
% ------------------------------
% CHAPTER: UBRAL
% ------------------------------
\chapter{Ubral}
\emph{The Stone Between Spears.}

\noindent
{\Large
\textit{The stone here remembers every oath sworn upon it and every blood spilled in its shadow. To lead is to understand that the land itself is the true sovereign, and we are merely its stewards.}

\vspace{0.5em}

I am a Hearth-Aunt. Hospitality's a sacred thing in these hills---break bread at my table and you're family till the sun sets twice. But cross that trust and the very stones will remember your name.

\vspace{0.5em}

Ubral is a land of upland mists, stone cairns, and clan oaths held tighter than iron. Caught between the Aelerian mountains and the lowland courts of Viterra, its hills are scarred by old raids and crowned by watch-fires that speak faster than any rider. Here, law is written not in charters but in cattle, wergild, and songs; a guest's word can buy more than a sword, but one broken vow can spark a feud lasting generations. But here's the secret the Lady of Tor doesn't want you to know: the cairns are not just markers. They are witnesses. Every stone in every cairn has a name. If you touch the right stone, you can hear the dead voting on the disputes of the living. The dead have long memories. They vote with the wind.

\vspace{0.5em}

Viterra has tried to conquer Ubral a dozen times; each attempt ended with supply lines cut by reivers, passes sealed by Aeler outcast engineers, and lowland knights lost in mist. The Utaran Imperium itself broke its leg on these hills and limped away. Now the clans watch the lowlands with patient eyes, and the exiled Aeler of Khaz-Vurim weigh every traveler's honesty on iron scales. A guest here is kin---until they prove otherwise. And kin here will die for you---or kill you over a stolen cow.

\vspace{0.5em}

Never refuse a cup from a hearth-aunt. Hospitality is a contract sealed in salt and bread. Refusal is an insult that blood must answer.

\vspace{1em}

\noindent\textit{--- Hearth-Aunt Brigid, who has never drawn a blade and has never needed to}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Oatcakes and skyr---the oats are grown in highland clearings, the skyr is cultured in stone crocks buried in the peat. The peat is ancient. The peat remembers the old kings. The old kings are buried beneath it.

\textbf{What does a child learn before they can walk?} To count cattle. Ubral children are taught the eight breeds of the hills: Black, Red, White, Brindled, Dun, Gray, Spotted, and the ninth---the breed that has no color because it belongs to the cairn-folk. The cairn-folk are the dead. The dead do not pay taxes.

\textbf{What is the first thing you notice about a stranger?} Their guest-token. Every Ubral traveler carries a token---a stone, a bone, a braid of hair---given by their last host. A stranger whose token is warm has been recently welcomed. A stranger whose token is cold has been traveling long. A stranger with no token is a spy. Spies are given a token---a rope. The rope is for their neck.

\textbf{What is the most common crime?} Cattle raiding---but raiding is not a crime, it is a sport. The cattle are the wealth. The wealth is the clan's. A stolen cow is a stolen child. The child must be returned. The return is a negotiation. The negotiation is a feud. The feud is a song.

\textbf{What does no one talk about but everyone knows?} The Lady of Tor is not the ruler. The cairns rule. The cairns vote at each full moon. Their votes are carried by the wind. The wind speaks to the hearth-aunts. The hearth-aunts speak to the clan. The clan obeys. The Lady is the voice. The cairns are the vote.

\textbf{Where do people go when they die?} To the cairns. The dead are not buried. Their bones are added to the cairns. The cairns grow taller each year. The tallest cairn is the oldest. The oldest cairn has no name. The oldest cairn remembers the first clan.

\textbf{How do you apologize for a serious wrong?} You offer a wergild---a payment in cattle, calculated by the wronged party's estimate of their loss. The cattle are driven to the cairn. The cairn is the court. If the cattle graze peacefully, you are forgiven. If they scatter, you are not. If they refuse to enter the cairn's shadow, your debt is blood.

\textbf{What is the most important festival?} The Gathering of the Clans (autumn). The clans meet at the Great Cairn. The cattle are counted. The songs are sung. The feuds are settled. The dead are named. The wind carries the names to the hills. The hills remember.

\textbf{What is the secret that could destroy a powerful person?} The Great Cairn is not a cairn. It is a prison---the tomb of a Viterran king who tried to conquer Ubral and failed. He was buried alive, with his crown and his sword. His ghost votes with the cairns. He votes for war. The cairns have not yet agreed.
\end{almanac}

\newpage
\input{../../generators/ubral.tex}
\newpage

% ------------------------------
% CHAPTER: SILKSTRAND
% ------------------------------
\chapter{Silkstrand}
\emph{City of Bridges \& Dyewater.}

\noindent
{\Large
\textit{In this city, every bridge is a contract and every current carries coin. The bargains struck in our shadows never quite fade from memory, and the waters remember every handshake.}

\vspace{0.5em}

I am a Bobbin-Runner. I can cross this city faster than any grown-up, and I know which planks creak and which merchants lie. The bridges remember every step, and the canals carry more secrets than fish.

\vspace{0.5em}

Silkstrand is a city strung across canals and arches, a place where every bridge is a contract and every current carries rumor. The silk looms never sleep, the Exchange never closes, and the Archivolt's notaries seal fates with a drop of wax. Merchants rise and fall with the tide of the market, bravos carve reputations on the planks of Three-Queens Bridge, and the city's underworld runs on whispers and borrowed keys. But here's the secret the Matron doesn't want you to know: the canals are not just water. They are memory. Every bribe, every broken promise, every cargo that was never declared---the canals remember. And sometimes, at low tide, the water shows you the faces of the merchants who drowned rather than pay their debts.

\vspace{0.5em}

The Matron rules from her palazzo stairs with velvet and iron, but the city itself belongs to silk, silver, and shadow. The Exchange Volatility Timer ticks with every major cargo seizure, every trade route cut, every merchant default. When it fills, a panic hits---the Exchange closes, and all outstanding contracts are voided unless witnessed by the Matron herself. The Matron has voided nine contracts this year. The ninth was her own.

\vspace{0.5em}

Never name a rival while crossing a bridge at night. The water remembers, and the curse you send may come back wrapped around your own throat.

\vspace{1em}

\noindent\textit{--- Bobbin-Runner, who ties a knot in her apron string for every secret she carries}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with dried fish---the rice is from the Oshiiran barges, the fish is from the Zakov fleets. Both are smuggled. Both are delicious. Both are paid for in favors.

\textbf{What does a child learn before they can walk?} To count bridges. Silkstrand children are taught the eight bridges of the city: Merchant's, Weaver's, Dyer's, Cooper's, Pilot's, Beggar's, Saint's, and the ninth---the bridge that is only there at low tide. The ninth bridge leads to the Sunken Quarter. The Sunken Quarter is where debts go to drown.

\textbf{What is the first thing you notice about a stranger?} Their ink. Every Silkstrand merchant has a seal-ink stain on their right thumb. A stranger whose ink is fresh is new to the trade. A stranger whose ink is faded has been retired. A stranger with no ink is not a merchant. Not a merchant is not a person.

\textbf{What is the most common crime?} Contract fraud---but fraud is not a crime, it is a negotiation. The contracts are written in disappearing ink. The ink's formula is a secret. The secret is held by the Archivolt notaries. The notaries sell the antidote. The antidote is the fee.

\textbf{What does no one talk about but everyone knows?} The Matron is not the ruler. The canals rule. The canals set the prices. The canals decide who lives and who drowns. The canals are not water. The canals are the city's memory. The memory is the Matron's. The Matron is the canals' voice.

\textbf{Where do people go when they die?} To the canals. The dead are not buried. Their bodies are wrapped in silk and sunk in the deep channels. The channels are the city's ledger. The ledger is never balanced.

\textbf{How do you apologize for a serious wrong?} You offer a bridge---a toll paid to the Bridge-Lords. The toll is calculated by the wronged party's estimate of their loss. The toll is paid in silver. The silver is melted and poured into the canal. The canal remembers the weight. The weight is your apology.

\textbf{What is the most important festival?} The Dyeing of the Waters (spring). The dyers empty their vats into the canals. The water turns red, blue, green, gold. The colors are the year's fashions. The fashions are set by the Matron. The Matron's favorite color is the one that sells best. The best color is the one she dyed her own hair.

\textbf{What is the secret that could destroy a powerful person?} The canals are not natural. They were dug by a Sidhi engineer who was paid in promises. The promises were broken. The engineer cursed the city. The curse is that every ninth drop of water is a tear. The tears belong to the engineer. She is still crying. She has been crying for nine centuries.
\end{almanac}

\newpage
\input{../../generators/silkstrand.tex}
\newpage

\chapter{Midh-Akhaz}
\emph{The Violet Gate.}

\noindent
{\Large
\textit{Coin talks. Grass walks. Blood settles the rest.}

\vspace{0.5em}

I am a Horse-Braid Apprentice. I tie a new knot for every paying passenger. My hair is a map of debts --- and grudges. The wind carries names, not ledgers, but here, in the shadow of the Oasis Compact, both trade and oaths change hands. Do not mistake our patience for weakness. The grass remembers every hoof that crossed it.

\vspace{0.5em}

Midh Ahkaz is not a city. It is a wound that learned to trade. A day's ride north of the Ecktorian border, it squats on the edge of the Violet Steppe like a patient wolf. No walls. No gates. Just rings of felt tents, latrine trenches, and the smell of horse sweat, cedar smoke, and the faint copper tang of blood money.

\vspace{0.5em}

The grass grows tall enough to hide a body. The locals cut it back at dawn --- but never at dusk, because the things that crawl through it at night have sharp teeth and no patience for property lines. Ecktorian emissaries call this place ``the Brink.'' Caravan guards call it home. Thieves call it paradise.

\vspace{0.5em}

Every morning, the Coin-Weigh Tribunal opens its scales under a canvas dome. Every night, the Dust Market auctions goods that fell off a wagon --- or off a corpse. The Khatun's word is law until someone stronger speaks. And in the Silver Stall Stable, a horse thief can buy a fresh mount with a story and a knife.

\vspace{0.5em}

Never refuse the guest-cup. The first pour is courtesy. The second is negotiation. The third is a blood-price unspoken. And the fourth --- the fourth means you're already dead, you just haven't stopped breathing.

\vspace{1em}

\noindent\textit{--- Horse-Braid Apprentice, who ties a knot for every secret she carries}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Curd cheese and flatbread --- the cheese is from the Ykrul herds, the bread is baked in the embers of a dung fire. It tastes of smoke and the open sky. The poor dip it in salt; the rich dip it in honey stolen from a border hive.

\textbf{What does a child learn before they can walk?} To count horses. Midh Ahkaz children are taught the eight gaits of a healthy mount: walk, trot, canter, gallop, pace, amble, rack, and the ninth --- the death-sprint that only a stolen horse knows. The ninth gait is never taught; it is discovered when the rider's life depends on it.

\textbf{What is the first thing you notice about a stranger?} Their braids. Ykrul warriors braid their hair with clan colors; Vilikari traders wear oiled braids with silver beads. A stranger whose braids are loose has just fought. A stranger whose braids are cut has lost a blood-price. A stranger with no braids is either a slave or a ghost. Neither stays long.

\textbf{What is the most common crime?} Grass-right forgery --- but forgery is not a crime, it is a negotiation. The Oasis Compact issues branded sticks for grazing rights. A clever carver can duplicate the brand. The penalty is exile into the tall grass at dusk. The grass does not have a court of appeals.

\textbf{What does no one talk about but everyone knows?} The Khatun is not the only ruler. The Silver Stall Stable master knows where every stolen horse sleeps. The Dust Market auctioneer knows the price of every sin. And the Iron Well keeper knows the name of everyone who will die before the next grass moon. The three of them meet once a season. No one records what they say. The grass records it.

\textbf{Where do people go when they die?} To the grass. The dead are not buried. Their bodies are carried beyond the caravan rings and left for the wolves and the wind. The grass grows tall and green where a body falls. A patch of deep violet is a sign that the dead were not properly mourned. The locals avoid those patches.

\textbf{How do you apologize for a serious wrong?} You offer a blood-price in silver, weighed on the Coin-Weigh Tribunal's scales. If the scale tips in your favor, you are forgiven. If it tips against you, you owe a horse. If the scale balances, you owe a story --- and the story must be true, and the true story must make the wronged party weep. If it does not, you owe a finger.

\textbf{What is the most important festival?} The Grass Moon (autumn equinox). The violet seed-heads glow under moonlight. The tribes ride in a great circle, singing the names of the dead. At midnight, the Khatun pours a cup of water onto the earth. If the water sinks, the winter will be mild. If it pools, the winter will be hungry. If it steams, the Hollow is walking. The festival ends at dawn. No one sleeps.

\textbf{What is the secret that could destroy a powerful person?} The Oasis Compact's original seal was not lost. It was stolen by a Vilikari factor who wanted to control water rights. The factor is still alive, living in the Foreign Quarter under a false name. The seal is hidden in a hollow horse skull buried under the Grass Ring. The skull belonged to the Khatun's favorite stallion. The stallion did not die of old age.
\end{almanac}

\newpage
\input{../../generators/midh-akhaz.tex}
\newpage

% ------------------------------
% CHAPTER: THE WAYS BETWEEN
% ------------------------------
\chapter{The Ways Between}
\emph{Spiritways \& Veilways.}

\noindent
{\Large
\textit{You are reading this sentence twice. The first time, you understood it. The second time, it understood you.}

\vspace{0.5em}

I am a Lost Pilgrim. Or I was. Or I will be. The paths here don't care about tenses. They care about which version of you arrives at the end. I've been walking for---no. That's a lie. The walking has been walking me. I forget which of us started first.

\vspace{0.5em}

The Ways Between do not measure miles. They measure choices. A wrong turn might lead to yesterday. A right turn might lead to a version of yourself who made a different decision and is still alive to regret it. The mist does not hide landmarks. It hides destinations. And sometimes, when you walk long enough, you realize the destination was never the point. The point was the moment you turned left instead of right. The point was the breath you held at the crossroads. The point was the name you did not speak.

\vspace{0.5em}

Here is the secret the Wayfinder doesn't want you to know: there is no Wayfinder. There is only the road, and the road is a liar. It tells you it wants you to find your way. What it actually wants is to watch you choose. Every reflection in every pool is a version of you that chose differently. They are all still walking. They are all still watching. And they are all still waiting for you to make the choice that finally lets them stop.

\vspace{0.5em}

I met myself on a bridge once. They were coming from the opposite direction. We stopped. We looked at each other. We said nothing. We both knew that if one of us spoke, the other would have to listen. And listening is a kind of debt.

\vspace{0.5em}

Never answer a question asked by a reflection. The answer will bind you to the path you did not choose. But here is the second secret: not answering is also a choice. And the road is very, very patient.

\vspace{1em}

\noindent\textit{--- The Wayfinder, who exists only in the moment a choice is made, and who is not the same person at the end of this sentence as at the beginning}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Nothing. There is no breakfast in the Ways Between. There is no morning. There is no hunger. There is only the walking. The walking is sustenance.

\textbf{What does a child learn before they can walk?} To forget. Children born in the Ways Between do not learn. They remember. They remember the path they took to be born. The path has eight turns. The ninth turn is the birth. The birth is a choice. The child does not know they chose.

\textbf{What is the first thing you notice about a stranger?} Their direction. A stranger walking toward you may be coming from your future. A stranger walking away may be going to your past. A stranger walking beside you may be yourself from a timeline where you did not turn left. Do not ask them which. They will not answer.

\textbf{What is the most common crime?} Stealing---but stealing is not a crime, it is a paradox. A thief in the Ways Between steals a memory from a version of you that no longer exists. The memory never belonged to you. You cannot miss it. You will miss it anyway.

\textbf{What does no one talk about but everyone knows?} The Ways Between are not between anything. They are the only thing that exists. The solid ground you remember---the cities, the roads, the hearths---is the dream. The Ways Between are the waking. The waking is the walking. The walking never ends.

\textbf{Where do people go when they die?} Deeper. The Ways Between have no surface, no bottom, no end. The dead walk deeper. The deeper they walk, the less they remember. When they remember nothing, they become the path. The path is the walking. The walking is the dead.

\textbf{How do you apologize for a serious wrong?} You cannot. Apologies require memory. Memory requires a self. The Ways Between erase both. The only apology is to stop walking. Stopping is not an option.

\textbf{What is the most important festival?} There are no festivals. There is only the moment between heartbeats when you realize you have been walking for nine years and you do not remember why. That moment is the festival. Celebrate it by taking the next step.

\textbf{What is the secret that could destroy a powerful person?} The Wayfinder is not a person. It is the road itself, dreaming of a traveler who will finally reach the end. The end does not exist. The road knows this. The road dreams anyway.
\end{almanac}

\newpage
\input{../../generators/ways_between.tex}
\newpage

% ------------------------------
% CHAPTER: THE WILDS
% ------------------------------
\chapter{The Wilds}
\emph{Roads, Ruins, and Weather.}

\noindent
{\Large
\textit{Between towns and treaties lies the patience of the land. Old roads remember armies; cairns remember names no book keeps. In the Wilds, law thins to trail-marks and favors, and the weather keeps its own counsel. The Edgewalker calls it ''unclaimed inventory.`` The land calls it patience. The land has more patience than the Edgewalker.}

\vspace{0.5em}

I am a ranger now. Stonehaven trained me as a scholar---mapping ruins, copying old ledgers, believing that ink could preserve what the world forgot. Tam taught me that the world does not forget. It waits. And while it waits, it watches. The wilds don't need your roads or your rules; they've got their own ways of keeping score. Follow the cairns if you want to live, and always leave something behind---a coin, a story, a breath held just long enough to be noticed. Every hill hides a secret, and every hollow breathes.

\vspace{0.5em}

The land has a ledger too. It just doesn't use ink. Between the cartographers' claims, the wilds endure---uninterested in borders, indifferent to tariffs. Here, rangers read spoor and weather; hermits keep shrines at forgotten fords; beasts and older things observe travelers with eyes that have seen empires rise and fall. Travel is a negotiation with the land itself, and the land always remembers a slight. The hollows have long memories. They are not cruel. They are not kind. They are patient.

\vspace{0.5em}

Never sleep in a hollow that has soot on the ceiling without adding your own to the fire. The land knows the difference between a guest and a thief. It also knows the difference between a thief and a corpse. It prefers the corpse. Less arguing.

\vspace{1em}

\noindent\textit{--- Moira of Stonehaven, ranger and reluctant debt-keeper}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Whatever they found---berries, roots, a rabbit snared at dawn. The wilds provide. The wilds also take. The taking is breakfast for something else.

\textbf{What does a child learn before they can walk?} To read the sky. Wilds children are taught the eight winds by taste: North is dry, South is wet, East is cold, West is warm, and the ninth---the wind that has no direction because it is the land's own breath. The land's breath is a warning.

\textbf{What is the first thing you notice about a stranger?} Their pack. A stranger whose pack is heavy has been traveling long. A stranger whose pack is light has just started. A stranger with no pack is either a fool or a ghost. The wilds eat fools. Ghosts eat the wilds.

\textbf{What is the most common crime?} Forgetting---but forgetting is not a crime, it is a survival strategy. The wilds are full of things that want to be forgotten. A traveler who forgets them may pass safely. A traveler who remembers must pay a toll. The toll is a story. The story is the memory. The memory is the trap.

\textbf{What does no one talk about but everyone knows?} The wilds are not empty. They are full of watchers---spirits, beasts, old things that do not have names because names are debts. The watchers do not interfere. They watch. They wait. They remember every footprint. When the footprints stop, the watchers move on.

\textbf{Where do people go when they die?} Into the land. The dead are not buried. They are left where they fell. The land absorbs them. The land grows a little richer. The land's memory grows a little longer. The dead become the land's memory. The memory is the wilds.

\textbf{How do you apologize for a serious wrong?} You leave an offering---a coin, a button, a lock of hair---at the place where the wrong was done. The land accepts or refuses. If the offering is gone by morning, you are forgiven. If it remains, you are not. If it has been moved, the land is testing you. The test is the walk. The walk is the apology.

\textbf{What is the most important festival?} The Turning of the Leaves (autumn). The wilds change color. The colors are the land's mood. Red is anger. Gold is warning. Brown is patience. Green is forgetfulness. This year, the leaves are red. The land is angry. Do not ask why.

\textbf{What is the secret that could destroy a powerful person?} The wilds have a heart. It is a stone, buried beneath the oldest oak. The stone is warm. The stone is beating. The stone is the land's pulse. If the stone stops, the wilds die. If the wilds die, there is nothing left to remember. Nothing left to wait.
\end{almanac}

\newpage
\input{../../generators/wilderness.tex}
\newpage

% ------------------------------
% CHAPTER: DUNGEONS
% ------------------------------
\chapter{Dungeons}
\emph{Living Infrastructure.}

\noindent
{\Large
\textit{Beneath the surface, ancient places breathe with mechanical lungs and dream electric dreams. The walls remember every footfall; the air carries whispers of forgotten conversations. The Edgewalker believes dungeons are infrastructure. The prisoners believe they are hungry. Both are correct.}

\vspace{0.5em}

I am a Scholar-Prisoner. The walls here have a memory like old parchment---they remember every scream, every spell, every drop of blood spilled on the stones. And sometimes, when the air's just right, they whisper it back to you. The dungeon listens. The question is whether it learns. I have been here for nine years. I have heard it learning. It started with single words. Then sentences. Now it speaks in paragraphs. Last week, it finished my thoughts before I did.

\vspace{0.5em}

Dungeons are not ruins---they are systems. Ventilation shafts that breathe, water channels that pulse, doors that open and close without any hand to move them. The Aeler built many of them; the Covenant sealed others; the Sidhi mined a few; and some were never built at all. They grew. Like fungus in the dark. Like a story told too many times. Like a debt that compounds.

\vspace{0.5em}

Never ignore a door that opens when you're not looking. The dungeon is offering you a choice. Refuse, and the next door may not open at all. Accept, and the door may lead to a room that was always there, waiting for someone to finally walk through. The room has been waiting for nine thousand years. It is not tired of waiting.

\vspace{1em}

\noindent\textit{--- The Dungeon's Architect, whose plans are written in a language no one speaks and whose signature is a thumbprint that could be anyone's}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Fungus bread and cave water---the fungus grows on the walls, the water drips from the ceiling. Both taste of stone. Both are all there is.

\textbf{What does a child learn before they can walk?} To count footsteps. Dungeon children are taught the eight sounds of the deep: Drip, Creak, Groan, Whisper, Scrape, Echo, Silence, and the ninth---the sound of the dungeon's own heartbeat. The heartbeat is a door opening. The door is always behind you.

\textbf{What is the first thing you notice about a stranger?} Their shadow. Dungeon shadows are long, even when the light is bright. A stranger whose shadow is too short has been here too long. A stranger whose shadow moves independently has been touched by the deep. A stranger with no shadow is not a stranger. They are the dungeon's guest. The dungeon never lets guests leave.

\textbf{What is the most common crime?} Map-making---but map-making is not a crime, it is a heresy. The dungeon does not want to be mapped. The dungeon changes its corridors when it feels watched. A map that is accurate at dawn is a lie by dusk. The lie is the map-maker's fault. The punishment is to walk the lie until it becomes true.

\textbf{What does no one talk about but everyone knows?} The dungeon is not a place. It is a creature---a sleeping thing that dreamed itself into stone. The corridors are its veins. The chambers are its organs. The prisoners are its thoughts. The thoughts are not its own.

\textbf{Where do people go when they die?} Into the walls. The dead are not buried. Their bones are ground into mortar and used to patch the cracks. The cracks always reappear. The dead are never finished being used.

\textbf{How do you apologize for a serious wrong?} You offer a memory---a true one, witnessed by the dungeon itself. The dungeon accepts the memory. The memory is added to the walls. The walls whisper it back to you when you sleep. The whisper is the apology. The apology never ends.

\textbf{What is the most important festival?} The Turning of the Keys (whenever the dungeon decides). The locks are changed. The prisoners are moved. The corridors are rearranged. The festival has no date, no warning, no end. It is happening now. You did not notice. That is the point.

\textbf{What is the secret that could destroy a powerful person?} The dungeon's architect is still alive. They are sealed in the lowest chamber, kept alive by the dungeon's own will. The dungeon does not want to be alone. The architect is the dungeon's only friend. The architect has not spoken in nine centuries. The dungeon does not need words.
\end{almanac}

\newpage
\input{../../generators/dungeons.tex}
\newpage

% ============================================================
% EPILOGUE: THE LEDGER CLOSES
% ============================================================
\chapter*{Epilogue: The Ledger Closes}
\addcontentsline{toc}{chapter}{Epilogue: The Ledger Closes}

\noindent
{\Large
I closed the book. The ink was dry.

\vspace{0.5em}

I had written the final entry months ago, in a damp counting-house on the Dhaharan coast, by candlelight. My hands did not shake---they were too good for that---but I set down my abacus very carefully. I had found something in the margins. A footnote I did not remember writing. A clause I did not remember adding. The ink was the same color as my own blood. The handwriting was my own. The date was tomorrow.

\vspace{0.5em}

I have learned to trust the things I do not remember. They are usually the truest things I know.

\vspace{0.5em}

The hat---the one the Trow had given me, the one that was still warm---sat on the corner of the desk. I picked it up, turned it over in my hands, and for the first time in years, I could not think of a single angle. That should have frightened me. Instead, it felt like a door opening. Not the kind of door you walk through. The kind of door that walks through you.

\vspace{0.5em}

The road was south. I did not know who would be waiting. I went anyway.
}

\vspace{1em}

\noindent\textit{--- The Edgewalker}

\vspace{1em}

% After the Edgewalker's final line, add this:

\vspace{2em}
\begin{quote}
\small
\textit{``The road is a ledger. Every oasis is a credit, every broken wheel a debt. The Tulkani tie knots to remember who owes whom. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. And the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.''}

\noindent --- Dusana of the Raven Road, at a caravanserai whose name she refused to give
\end{quote}