\part{The Way of Silk}

% After \part{The Way of Silk}
\begin{quote}
  \small
  \textit{``The road is a ledger. Every oasis is a credit, every broken wheel a debt. The Tulkani tie knots to remember who owes whom. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. And the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.''}
  \noindent --- Dusana of the Raven Road, at a caravanserai whose name she refused to give
  \end{quote}
% ------------------------------
% CHAPTER: FHARA
% ------------------------------
\chapter{Fhara}
\emph{The Oasis Courts.}

\noindent
{\Large
\textit{Water is not a right. It is a loan from the sand. The oasis remembers every drop drawn, every cup poured, every promise made at its edge. I do not judge the living. I judge the debt.}

\vspace{0.5em}

I am a Water-Clerk. My grandfather dug this well with a shovel and a prayer. My mother sealed the walls with clay from the dry river. I sit in the shade and measure every bucket. The sand does not forgive a miscount.

\vspace{0.5em}

Water isn't just law in Fhara---it's a market, and oh, how the water-clerks love a good deal on aquifer access. They don't just negotiate in three cups; they auction the right to the fourth cup. The Fhara are merchants, poets, and negotiators; they will bargain for an hour over a cup of coffee, and a day over a camel. But here's the secret the well-judges don't want you to know: the wells have ears. Every cup poured, every promise made at the edge, every drop drawn---the well-spirits are listening. And they have opinions. They favor the generous. They punish the greedy. And they never, ever forget a face.

\vspace{0.5em}

The Fhara have no single king. Their confederacy is a web of marriage alliances, water-sharing treaties, and caravan tariffs. The desert is not empty---it is a ledger of dry wells, abandoned caravanserais, and the bones of those who could not share. A Fhara merchant can cross the waste with only a waterskin and a coffee pot, because every well has a keeper, every keeper a price, and every price can be negotiated. But the negotiation is a ritual. Three cups: the first is welcome, the second is bargaining, the third is a contract. Refuse the third, and you have broken an oath. The well-spirit will remember.

\vspace{0.5em}

Never refuse a date offered by a well-keeper. The fruit is sacred to their well-spirits. Refusing it ends the conversation---and your access to water.

\vspace{1em}

\noindent\textit{--- Well-Judge Fatima al-Wadi, who carries a dipper made from a date-palm leaf that rots after each case}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Dates and thin coffee---the dates are from the well-keepers' groves, the coffee is brewed with cardamom and camel milk. The milk is sour. The sourness is for endurance.

\textbf{What does a child learn before they can walk?} To count cups. Fhara children are taught the eight cups of negotiation: Welcome, Bargain, Contract, Witness, Debt, Payment, Forgiveness, and the ninth---the cup that is never poured. The ninth cup is for the well-spirit. The well-spirit is always thirsty.

\textbf{What is the first thing you notice about a stranger?} Their water-skin. A stranger whose water-skin is full has not traveled far. A stranger whose water-skin is dry has traveled too far. A stranger who carries no water-skin is either a well-keeper or a fool. Well-keepers are honoured. Fools are buried.

\textbf{What is the most common crime?} Water theft---but water theft is not a crime, it is a declaration of feud. The well-spirit records every theft in the aquifer. The aquifer is the court. The judge is the well. The verdict is thirst.

\textbf{What does no one talk about but everyone knows?} The wells are not natural. They were carved by the first Fhara, who made a bargain with the sand. The bargain was for water. The price was memory. The Fhara remember everything. They cannot forget. That is the curse of the wells.

\textbf{Where do people go when they die?} To the oasis. The dead are buried in the date groves, their graves marked with a single stone. The stone is not carved. The dead's name is spoken at sunrise. If the name is remembered, the dead rests. If it is forgotten, the dead walks.

\textbf{How do you apologize for a serious wrong?} You offer a cup---the ninth cup, poured from your own water-skin. The cup is placed at the well's edge. If the well-spirit drinks, you are forgiven. If the water evaporates, you are not. If the cup cracks, your debt is doubled.

\textbf{What is the most important festival?} The Filling of the Wells (spring). Every well is cleaned and blessed. The water-clerks recite the names of the year's debtors. The debtors stand at the water's edge. If their reflection wavers, they are forgiven. If it holds, they must pay.

\textbf{What is the secret that could destroy a powerful person?} The water-clerks' ledger is a forgery. The original was lost in a sandstorm three generations ago. The clerks have been inventing water rights ever since. The wells do not care. The wells remember the original. The wells are waiting.
\end{almanac}

\newpage
\input{../../generators/fhara.tex}
\newpage

% ------------------------------
% CHAPTER: SIDHI
% ------------------------------
\chapter{Sidhi}
\emph{The Diaspora of the Sealed Coast.}

\noindent
{\Large
\textit{Galanina is not a place on a map. It is a promise that our ancestors made to the sea: we will remember, and we will return. The Covenant gave us courts, but the sea gave us memory. The Admiralty gives us tariffs. We give each other salt and stories.}

\vspace{0.5em}

I am a Freedman Sailor. My grandmother wore a brass collar. My mother wore a freedman's seal. I wear a captain's ring. The slavers took our names, but they could not take the tide. The tide always comes back. So do we.

\vspace{0.5em}

The Sidhi are a people of the coast and the diaspora. Their ancestral homeland is Galanina, a rugged shore of limestone cliffs and ancient ports far to the west. The Ashaani magocracy enslaved them for generations, but Galanina broke free under the Covenant of Mykkiel. Now the Sidhi are scattered, fierce, proud, and never forget a debt. But here's the secret the Admiralty doesn't want you to know: the Sidhi never stopped running. The Long Trail---the forced migration that scattered them across the Amaranthine---is not a tale told to children. It is a silence. A pause in conversation when someone mentions the eastern passes or the southern coast. An old woman who stops mid-sentence, touches her brass key, and changes the subject. The Sidhi do not forget, but they have learned that memory is a kind of weight---and they have carried it far enough.

\vspace{0.5em}

The Sidhi of the Brass Coast run the ropeways, the spice auctions, and the port guilds. Their speech is a creole of Ashaani and coastal tongues, their law is maritime, and their symbol is the lighthouse lamp---a green flame that shows the faces of the drowned, so that the dead may witness the living. They are welcome everywhere and trusted nowhere. They prefer it that way.

\vspace{0.5em}

Never ask about ancestry. The Ashaani slavery raids are a fresh wound. Asking which clan someone belongs to is asking how their grandparents were captured, sold, or killed.

\vspace{1em}

\noindent\textit{--- Captain Idris, Sidhi lighthouse commander, whose left eye is a brass lens}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salted fish and flatbread---the fish is dried on the docks, the bread is baked in communal ovens. The ovens are fuelled by driftwood. The driftwood carries the names of lost ships.

\textbf{What does a child learn before they can walk?} To read the lighthouse codes. Sidhi children are taught the eight lamp-flashes: Safe, Danger, Storm, Pirate, Customs, Quarantine, Home, and the ninth---the flash that signals a freedman's return. The ninth flash is never used. The ninth flash is a promise.

\textbf{What is the first thing you notice about a stranger?} Their seal. Every Sidhi freedman carries a brass seal---the mark of their manumission. A stranger whose seal is worn smooth has been free for generations. A stranger whose seal is sharp has just been freed. A stranger with no seal is not a Sidhi. Not a Sidhi is not a person.

\textbf{What is the most common crime?} Seal-forging---but seal-forging is not a crime, it is a necessity. The Admiralty demands proof of freedom. The proof can be forged. The forgers are Sidhi. The Sidhi do not prosecute forgers. The Sidhi hire them.

\textbf{What does no one talk about but everyone knows?} The brass seals are not brass. They are made from melted chains---the chains that bound the first freedmen. The metal remembers the weight of slavery. The seals are warm. The warmth is the memory of wrists.

\textbf{Where do people go when they die?} To the lighthouse. The dead are not buried. Their ashes are scattered from the tower's top. The wind carries them to the sea. The sea carries them to Galanina. Galanina is the shore of ancestors. The ancestors are waiting.

\textbf{How do you apologize for a serious wrong?} You offer your seal---your own mark of freedom. The seal is melted down and reforged into a chain. The chain is worn by the wronged party. When the wrong is forgiven, the chain is melted again. The metal grows thinner each time. One day, there will be nothing left.

\textbf{What is the most important festival?} The Lighting of the Green Flame (winter solstice). The lighthouse lamps are cleaned and refilled. The green flame is lit. The flame shows the faces of the drowned. The drowned are the ancestors. The ancestors watch. The ancestors do not judge.

\textbf{What is the secret that could destroy a powerful person?} The brass seals are not all equal. Some were forged by the Admiralty itself---trap-seals that mark the wearer as a spy. The spies do not know they are spies. Their reports go to the Admiralty. The Admiralty knows everything. The Admiralty has been playing a very long game.
\end{almanac}

\newpage
\input{../../generators/sidhi.tex}
\newpage

% ------------------------------
% CHAPTER: PERESHI
% ------------------------------
\chapter{Pereshi}
\emph{The Roof of the Way.}

\noindent
{\Large
\textit{The fire does not forget. It has witnessed every oath sworn in this valley, every treaty sealed with a dropped coal, every lie told in its light. I do not tend the flame. I listen to what it remembers. And it remembers everything.}

\vspace{0.5em}

I am a Caravan Scribe. My grandfather copied a tax ledger that had not been opened in three centuries. He found a mistake---a single miscounted bale of silk. That mistake started a war. We do not write for the living. We write for the dead who cannot correct their errors.

\vspace{0.5em}

The Pereshi highlands are a vast, windswept plateau, cut by deep river valleys and crowned with snow-capped peaks. Ancient fire temples stand beside caravanserais; terraced vineyards climb the hillsides. The people are scholars, shepherds, and traders---proud, suspicious, and bound by a code of hospitality that demands three days of feasting for any guest, and three generations of vengeance for any insult. But here's the secret the Flame-Keepers don't want you to know: the fire in the temple is not just a flame. It is a witness. It has seen every oath sworn in this valley, every treaty sealed with a dropped coal, every lie told in its light. And it remembers everything. Including the Prince's true name.

\vspace{0.5em}

Once, the Forgotten Ashaani Prince conquered this land. He built garrisons, taxed the passes, and tried to burn every library. When he and his kingdom vanished into the sand---swallowed by the same curse that turned the desert to glass---the Pereshi rose. They threw down his garrisons, brick by brick, and sealed his name in a scroll they have never reopened. They have not forgotten. Every fire temple keeps a coal from the Prince's own brazier---the ninth coal, never lit, a promise that his darkness will not return.

\vspace{0.5em}

Never step on a fallen book. Written words carry the weight of law. To step on a book is to trample a contract, a genealogy, a verdict. The spirits of the words will haunt you.

\vspace{1em}

\noindent\textit{--- Flame-Keeper Zarifa, who has never been seen without a rose---fresh or dried}
}

\vspace{1em}

\begin{almanac}
  \textbf{What do people eat for breakfast?} Flatbread with honey and clotted cream---the honey is from highland bees, the cream is from goats that graze the temple steps. The goats are sacred. The goats are also delicious. The bread is baked in ovens heated with coals from the temple fire; families take turns carrying the embers home. A coal that goes out before dawn is a bad omen.
  
  \textbf{What does a child learn before they can walk?} To read a contract. Pereshi children are taught the eight clauses of the Covenant: Offer, Acceptance, Witness, Consideration, Term, Breach, Remedy, and the ninth---the silence where the unspoken terms live. The ninth clause is never written. The ninth clause is the one that kills you. Children also learn to check the ninth coal in the family shrine each morning. If it has grown warm overnight, they must wake the Flame-Keeper.
  
  \textbf{What is the first thing you notice about a stranger?} Their book. Every Pereshi carries a ledger---their personal account of debts owed and collected. A stranger whose ledger is thick has many debts. A stranger whose ledger is thin has many creditors. A stranger with no ledger is a liar. Liars are not fed. But these days, the first thing many notice is whether the stranger casts a shadow. The Prince's shadow has been seen walking alone.
  
  \textbf{What is the most common crime?} Contract forgery---but forgery is not a crime, it is an art. The Pereshi courts are full of forged documents. The judges are forgers themselves. The law is a negotiation. The negotiation never ends. The newest crime is shadow-theft: someone steals another's shadow to feed to the Prince. The victim becomes pale, listless, and cannot lie---but also cannot keep a secret.
  
  \textbf{What does no one talk about but everyone knows?} The Forgotten Prince did not vanish. He is buried beneath the largest fire temple, in a lead-lined coffin sealed with nine locks. The keys are held by the nine Flame-Keepers. The Flame-Keepers have never met. The Flame-Keepers do not know each other's names. The Prince's name is written on the inside of the coffin lid. No one has read it in nine centuries---but lately the lid has begun to sweat. A rust-colored liquid seeps from the seams.
  
  \textbf{Where do people go when they die?} To the fire. The dead are cremated on the temple pyres. Their ashes are mixed with clay and pressed into bricks. The bricks are used to build new temples. The dead become walls. The walls witness. The walls do not forget---and they whisper the names of those who still live. Some walls have started to whisper a name that is not on any ledger.
  
  \textbf{How do you apologize for a serious wrong?} You offer a coal---a burning ember from the temple fire. The coal is placed in the wronged party's hearth. If it stays lit, you are forgiven. If it smoulders, you are watched. If it goes out, you are cursed. The curse is the fire's silence. The fire will not speak for you again. Some say that if the coal goes out, the Prince's shadow will visit you that night. It asks only one question: ''Why should I not take your name?``
  
  \textbf{What is the most important festival?} The Reading of the Ledgers (autumn). Every household brings its ledger to the temple. The ledgers are read aloud. The debts are tallied. The dead are named. The fire listens. The fire does not interrupt---but the ninth coal flickers when a lie is spoken. This year, the coal has not stopped flickering.
  
  \textbf{What is the secret that could destroy a powerful person?} The ninth coal is not coal. It is a piece of the Forgotten Prince's heart, petrified by the curse that swallowed his kingdom. The Flame-Keepers do not know. They think it is a relic. The heart is still warm. The heart is still beating. The heart is waiting to be reunited with the rest of him. And someone has been feeding it blood---a drop each night, from a different Flame-Keeper. No one knows who. No one wants to know who.
  \end{almanac}

\newpage
\input{../../generators/pereshi.tex}
\newpage

% ------------------------------
% CHAPTER: KUVANI
% ------------------------------
\chapter{Kuvani}
\emph{The Remount Steppe.}

\noindent
{\Large
\textit{The grass does not belong to us. We belong to the grass. It tells us where to ride, when to move, how long to stay. A clan that forgets the grass's song is a clan that will be buried by it. I do not sing to the wind. I listen. The wind sings through me.}

\vspace{0.5em}

I am a Remount Master. A horse's name is its lineage. Its lineage is its debt. You do not ride a Kuvani horse. You borrow it. And when you return it, you will tell me a story that the horse has not already heard. If you lie, the horse will throw you next time.

\vspace{0.5em}

Horses, grass, and wind-songs---the Kuvani don't just pay in stories; they auction the right to the next remount. They range across the eastern steppe, a vast grassland east of the Valewood. Their horsemanship and remount stations make them indispensable to any caravan that crosses the open plain. The Kuvani have no cities; their courts are circles of saddles, and their laws are sung. But here's the secret the Wind-Singers don't want you to know: the sky is not empty. Kuva, the sky father, sees everything---every oath sworn in the open, every horse born, every blade of grass bent by the wind. But Kuva does not speak. He listens. And he is patient. The Kuvani have no temples because the sky is their shrine. Their dead are not buried; they are given to the sky on funeral platforms, that Kuva may carry their souls to the ancestral herd that rides the northern lights.

\vspace{0.5em}

A horse is not a mount---it is a relative. Its name is a lineage stretching back to the first mare that Kuva breathed life into. A Kuvani who cannot name his horse's ancestors is a Kuvani without a shadow. The steppe is not empty; it is full of ghosts: the ghost herd of the salt pan, the sky-road markers that hum at dusk, the Iron Corral where blades whisper the names of the dead.

\vspace{0.5em}

Never ride a Kuvani horse without asking its name. The horse will answer---by throwing you. If it throws you, you have insulted its lineage.

\vspace{1em}

\noindent\textit{--- Alp Arslan, Remount Master, who has never been seen sleeping}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Fermented mare's milk and dried meat---the milk is from the herd, the meat is from the hunt. The hunt is a ceremony. The ceremony is a debt to the grass.

\textbf{What does a child learn before they can walk?} To ride. Kuvani children are placed on a horse before they can stand. The horse is patient. The horse is old. The horse remembers the child's grandfather. The grandfather taught the horse to ride.

\textbf{What is the first thing you notice about a stranger?} Their saddle. A stranger whose saddle is worn has ridden many miles. A stranger whose saddle is new has many miles ahead. A stranger who carries no saddle is a ghost. Ghosts do not need saddles.

\textbf{What is the most common crime?} Horse theft---but horse theft is not a crime, it is a declaration of feud. The horse is family. A stolen horse is a stolen relative. The feud lasts until the horse is returned or the thief's line is ended. The feud never ends.

\textbf{What does no one talk about but everyone knows?} The ghost herd is not a legend. It is real---a herd of spectral horses that rides the salt pan on moonless nights. The horses are the souls of riders who died without honor. They ride forever, searching for a battle that will never come.

\textbf{Where do people go when they die?} To the sky. The dead are placed on funeral platforms, wrapped in felt. The birds take the flesh. The wind takes the bones. The sky takes the soul. The soul joins the ancestral herd. The herd rides the northern lights.

\textbf{How do you apologize for a serious wrong?} You offer a horse---the best horse in your herd. The horse is given to the wronged party. If they accept, the debt is paid. If they refuse, the debt is doubled. If they kill the horse, the debt is blood. Blood is never repaid.

\textbf{What is the most important festival?} The Ride of the Ancestors (autumn equinox). The clans ride to the salt pan. The ghost herd appears at midnight. The living ride with the dead. The dead remember their names. The living remember theirs.

\textbf{What is the secret that could destroy a powerful person?} The Khagan's horse is a ghost. It was killed in battle twenty years ago. The Khagan does not know. He rides a skeleton wrapped in hide. The horse does not tire. The horse does not eat. The horse does not forgive.
\end{almanac}

\newpage
\input{../../generators/kuvani.tex}
\newpage

% ------------------------------
% CHAPTER: TULKANI
% ------------------------------
\chapter{Tulkani}
\emph{The Itinerant Courts.}

\noindent
{\Large
\textit{The road does not forgive a debt. It only reschedules it. A promise spoken under the open sky is witnessed by the wind, the grass, and the wheels of every cart that passes. I do not write judgments. I tie knots. The knots remember. When you are ready to pay, untie one. If you are not ready, tie another.}

\vspace{0.5em}

I am a Wagon-Keeper. My wagon is my hearth, my court, my grave. The Hollow cannot cross the circle of wheels unless invited. The tax collector cannot enter without a story. The lord cannot demand rent unless we camp on his land for nine nights. On the eighth night, we leave. That is the law.

\vspace{0.5em}

The Tulkani have no fixed homeland; they are wanderers who follow the seasons and the trade routes. Their wagons are their hearths, their circles are their courts, and their knots are their ledgers. They are welcome everywhere and trusted nowhere---a position they wear with wry humour. They spread the cult of Ikasha---She Who Sleeps---not through temples, but through dream-cords tied in the dark. But here's the secret the Caravan Judges don't want you to know: the knots are not just records. They are alive. A knot tied in false witness will tighten until blood is drawn. A cord cut in anger will become a ghost that follows the oath-breaker until the promise is fulfilled. And the ninth knot---the one that is never tied---is the space where the Hollow waits.

\vspace{0.5em}

The Tulkani have no king, no judges, no written law---only the Caravan Circle, where disputes are settled by storytelling. A liar is exposed not by evidence but by the knots on his cord. Ikasha's cult is not a religion; it is a survival tactic---the dream inside the wagon, the whisper in the creaking wheel. The Tulkani pay no permanent taxes, but they pay tolls in stories. A lord who demands coin will be told a story so boring that his crops fail. A lord who offers hospitality will be told a story so beautiful that his harvest doubles. The Tulkani do not track debts; they track reciprocity. What you give, you get back---maybe tomorrow, maybe in the next valley. But the circle remembers.

\vspace{0.5em}

Never sleep outside the circle at night. The wagon ring is sanctuary. Outside it, the Hollow walks. A Tulkani will not rescue you---they will only mourn you at dawn.

\vspace{1em}

\noindent\textit{--- Elder Sarnai, Caravan Judge of the Nine Knots, whose cord drags on the ground}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Tea and hardtack---the tea is brewed from leaves gathered along the road, the hardtack is baked at the start of the season and lasts until the snow. The hardtack is always stale. The tea is always fresh. The contradiction is a lesson.

\textbf{What does a child learn before they can walk?} To tie a knot. Tulkani children are taught the eight journey-knots: Departure, Arrival, Promise, Debt, Payment, Witness, Apology, and the ninth---the knot that is never tied. The ninth knot is for the Hollow. The Hollow is not a creditor. The Hollow is a reminder.

\textbf{What is the first thing you notice about a stranger?} Their cord. Every Tulkani carries a knotted cord---their life's ledger. A stranger whose cord has many knots has traveled far. A stranger whose cord has few knots has just begun. A stranger with no cord is a ghost. Ghosts do not travel. Ghosts wait.

\textbf{What is the most common crime?} Story-theft---but story-theft is not a crime, it is a compliment. A stolen story is a story worth telling. The original teller is honoured. The thief is welcomed. The story grows with each telling. The story never dies.

\textbf{What does no one talk about but everyone knows?} The Tulkani are not wanderers by choice. They were exiled from their homeland nine centuries ago, for a crime that no one remembers. The crime is written in the ninth knot of the Elder's cord. The Elder's cord is too long to read. The Elder's cord drags on the ground.

\textbf{Where do people go when they die?} To the road. The dead are not buried. Their bodies are left at a crossroads. The road takes them. The road carries them to the next valley, the next camp, the next telling. The dead become stories. Stories are immortal.

\textbf{How do you apologize for a serious wrong?} You untie a knot---one knot from your cord, chosen by the wronged party. The knot is burned. The ash is scattered on the wind. If the wind blows toward the wronged party, you are forgiven. If it blows away, you are not. If it is still, the Hollow has accepted your apology. The Hollow's acceptance is a warning.

\textbf{What is the most important festival?} The Tying of the New Cord (new year). Every Tulkani receives a new cord, unknotted. The old cord is burned. The ash is mixed with water and drunk. The drink tastes of memory. The memory is the year that has passed. The year is never forgotten.

\textbf{What is the secret that could destroy a powerful person?} The Tulkani are not wanderers. They are pilgrims---walking toward a destination that no longer exists. Their homeland was swallowed by the desert. The desert is still hungry. The Tulkani are still walking. They will never arrive.
\end{almanac}

\newpage
\input{../../generators/tulkani.tex}
\newpage