\section{Rites of the Threshold}

These Rites are available to Runic Warriors sworn to Patrons of the Threshold family: The Sealed Gate, The Traveler, and The Pale Shepherd. They emphasize warding, passage, guidance, and the liminal spaces between.

\subsection{The Sealed Gate — Warden's Rites}

\subsubsection{Low Rite: Rite of the Sealed Threshold (4 XP)}

\textbf{Tags:} [WARD], [BIND], [THRESHOLD]

\textbf{Threshold:} A doorway, gate, or bridge.

Mark a doorway or passage as sealed. Unauthorized crossing tests Spirit + Resolve (DV 3) or suffers a Position penalty (one step worse for that crossing). Creates a 4-segment Ward Integrity clock.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Key's Rebuke (5 XP)}

\textbf{Tags:} [FORCE], [BANISH], [THRESHOLD]

\textbf{Threshold:} A barrier (door, gate, bridge) and a target attempting to cross.

The target tests Body + Resolve (DV 3) or is pushed back from the threshold. On a success, create a \textit{Rebuke Echo} token granting +1 die to the next boundary defense.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Circle of Denial (8 XP)}

\textbf{Tags:} [WARD], [AREA], [BANISH], [THRESHOLD]

\textbf{Threshold:} A salt ring, iron filings, or boundary stones.

Create a protective circle. Supernatural beings test against the caster's Spirit + Arcana (DV 4) to cross. Mortals suffer -2 dice to force entry. Creates a 6-segment Circle Integrity clock.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Writ of Passage (7 XP)}

\textbf{Tags:} [BOND], [COMMAND], [MOVEMENT], [THRESHOLD]

\textbf{Threshold:} Signed authorization or spoken permissions.

Designate an authorized passage route. Allies gain Position +1 when using the route. Unauthorized intruders suffer -1 die to movement or stealth. Creates an 8-segment Writ Authority clock.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Consecrated Barrier (13 XP)}

\textbf{Tags:} [WARD], [AREA], [SANCTUARY], [FATE], [THRESHOLD]

\textbf{Threshold:} Relics from three traditions, boundary markers, and a trespasser's acknowledgment.

Consecrate an area against all unauthorized passage. The tests required scale with the intrusion attempt's severity. A 10-segment Sacred Boundary clock tracks the barrier's strength.

\textbf{Obligation:} 3

\subsection{The Traveler — Road-Warden's Rites}

\subsubsection{Low Rite: Rite of Road-Sense (4 XP)}

\textbf{Tags:} [NAVIGATION], [AWARENESS], [TRAVEL]

\textbf{Threshold:} A road-nail or waystone pebble.

Unerringly pick the fastest safe route in Near or Far. Gain +1 die to avoid ambushes or delays. Create a 4-segment Path Memory clock to ignore difficult terrain once.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Traveler's Boon (5 XP)}

\textbf{Tags:} [MOVEMENT], [BOND], [HASTE], [TRAVEL]

\textbf{Threshold:} Thread tied at the wrist.

Ignore one level of difficult terrain or bureaucracy; +1 Effect to travel or escape checks. If shared, create a 2-segment Shared Journey bond with an ally.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Waymark (8 XP)}

\textbf{Tags:} [MARK], [GUIDANCE], [PATH], [TRAVEL]

\textbf{Threshold:} A chalk mark or small cairn.

Declare a lane as permitted or easy. Allies gain improved Position/Effect or ignore one obstacle. Create a 6-segment Marked Path clock.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Bridge Between (7 XP)}

\textbf{Tags:} [TELEPORT], [PATH], [BOND], [TRAVEL]

\textbf{Threshold:} Two pinches of road-dust clapped together.

Relocate a willing target within Far along a visible or named route. Unwilling targets may resist (Body + Resolve DV 3). Create a 4-segment Pathway Established clock.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Wayfarer's Covenant (14 XP)}

\textbf{Tags:} [VOW], [COMMAND], [BOND], [FEAR], [TRAVEL]

\textbf{Threshold:} Waystones from multiple regions.

Bind present parties to safe passage. While honored: +2 Effect on travel, reroll one failed travel roll per scene. Breaking the oath inflicts Harm 1 (Fatigue) and marks the breaker as \textit{Oathbreaker of the Road} (-2 dice to travel rolls).

\textbf{Obligation:} 3

\subsection{The Pale Shepherd — Death-Warden's Rites}

\subsubsection{Low Rite: Rite of the Gentle Guide (5 XP)}

\textbf{Tags:} [GUIDANCE], [BOOST], [THRESHOLD], [RISK]

\textbf{Threshold:} A small stone from a grave or threshold, warmed in the palm.

Choose one: ease a hard crossing (target gains +1 die to rolls involving change, decision, or emotional steadiness this scene) or find a hidden path (gain +2 dice to Navigation or Investigation when seeking unconventional routes).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Unburdened Farewell (6 XP)}

\textbf{Tags:} [HEAL], [CLEANSE], [DEBT], [FEAR]

\textbf{Threshold:} A personal letter never sent, burned in a copper dish.

Speak a final goodbye to someone not present—a lost lover, a dead parent, a friend you wronged. Clear 1 Fatigue and remove one psychological Condition (e.g., Haunted, Guilty) caused by that relationship. The GM gains 1 SB as the words drift into the threshold.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Passing Flock (8 XP)}

\textbf{Tags:} [GUIDANCE], [KNOWLEDGE], [HEAL], [AREA], [RISK]

\textbf{Threshold:} Black wool and a personal token of the one to be guided.

Choose one: open a brief passage through shadow-realms (target bypasses one insurmountable obstacle) or call shadow-sheep that carry fragments of weighty moments (retrieve lost details, soothe grief, or glean counsel from what was).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Steadfast Witness (7 XP)}

\textbf{Tags:} [WITNESS], [DEBT], [TRUTH], [THRESHOLD]

\textbf{Threshold:} A dying person's last words, spoken aloud and witnessed.

Record the dying's final testament. The words are binding on the living. Any who break them suffer -1 die to all actions for one week unless they atone. The GM gains 1 SB (the threshold remembers).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Unburdened Crossing (13 XP)}

\textbf{Tags:} [DEATH], [GUIDANCE], [FATE], [THRESHOLD]

\textbf{Threshold:} A dying person, a witness, and a grey wool cord.

Guide a willing creature across the threshold of death. They die peacefully, without fear. Before they pass, they may speak one final truth. That truth is considered witnessed; it cannot be denied by any who hear it. You gain a permanent memory of their last words, which may be used as a \textit{Truth Compass} (+1 die to any roll directly related to fulfilling their final wish).

\textbf{Obligation:} 3

\subsection{Rite Synergies}

Rites from different Threshold Patrons work well together:

\begin{itemize}
\item \textbf{The Sealed Gate + The Traveler:} Ward a route, then guide allies through it. Enemies are blocked; allies are sped.
\item \textbf{The Pale Shepherd + The Sealed Gate:} Seal a deathbed against interruption. The dying pass in peace, witnessed only by the Shepherd and the Gate.
\item \textbf{The Traveler + The Pale Shepherd:} Guide a soul across the final threshold. The road is long. The Shepherd walks it with them.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
