\section{The Pale Shepherd — Passage and Melancholic Wisdom}

\subsection{Domain and Epithets}

The Pale Shepherd walks the seam between all things—life and death, joy and sorrow, memory and forgetting, the known and the unmarked. Neither wholly kind nor cruel, the Shepherd guides those who must pass through the hardest doors. Cloaked, with shadow-sheep that bear the shed wool of old possibilities, the Shepherd offers safe conduct through liminal space for a price paid in memory, feeling, or fate.

The Shepherd has no will, only a path; no demand, only a presence. It does not seek followers, but those who stand at thresholds—of death, of grief, of irreversible change—may find the Shepherd already there, waiting to accompany them. Its demands are not commands but the natural toll of passage: a memory left behind, a name forgotten, a tear shed for no reason.

\subsection{Cultural Origins}

The Pale Shepherd appears wherever thresholds are crossed:

\begin{itemize}
\item In \textbf{Aelaerem}, the Shepherd is the Pale Gardener, who walks the hedgerows at twilight, collecting the souls of those who die far from home.
\item In \textbf{Aeler}, the Shepherd is the Breath-Accountant, who weighs the last exhalation of the dying.
\item In \textbf{Linn}, the Shepherd is the Fog-Walker, appearing on the horizon before a storm.
\item In \textbf{Vhasia}, the Shepherd is the Silent Notary, witnessing deathbed testaments.
\end{itemize}

Wherever mortals face the hardest thresholds, the Pale Shepherd is there—not to judge, not to command, but to accompany.

\subsection{Warrior Archetype: The Death-Warden}

A Runic Warrior of the Pale Shepherd is not a sentinel or a road-guide. They are a \textit{Death-Warden}—a warrior who stands at the threshold between life and death, guiding those who cross and protecting those who remain. Their Codex is a grey cloak or a shepherd's crook. Their Thiasos is a spirit-sheep that carries memories, a grey bird that sings at funerals, or a sentient bell that tolls once for each soul guided home.

\subsection{Codex Form}

The Death-Warden's Codex is a symbol of passage:

\begin{itemize}
\item \textbf{Grey Cloak:} A heavy woollen cloak, grey as twilight. It is warm even in winter. It smells of rain and old stone.
\item \textbf{Shepherd's Crook:} A wooden staff curved at the top. It can pull a willing soul from the threshold of death—once per patient.
\item \textbf{Memory Cord:} A braided cord of grey wool, with eight knots. The ninth is never tied. Each knot holds a memory of someone who has passed.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Spirit-Sheep:} A small grey sheep with eyes like river stones. It carries the shed wool of old possibilities. It can absorb a memory and hold it safely for one day.
\item \textbf{Grey Bird:} A small bird, often a sparrow or dove. It sings when someone is about to die. It sings when someone is born. It does not sing for anything else.
\item \textbf{The Grey Bell (Sentient):} A small bronze bell that hums, never rings. It speaks in the voice of the last soul it guided. It asks only to be carried to thresholds.
\end{itemize}

\subsection{Core Mechanic: The Threshold Guide}

When the Death-Warden is present at a death (their own or another's), they may choose to guide the dying. The dying passes peacefully, without fear. The Death-Warden gains 1 Boon, which may only be used to ease suffering, protect the vulnerable, or cross a threshold.

When the Death-Warden helps someone through a difficult passage (grief, change, loss), they may clear 1 Fatigue. This can be done once per scene.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Gentle Guide (5 XP)}

\textbf{Tags:} [GUIDANCE], [BOOST], [THRESHOLD], [RISK]

\textbf{Threshold:} A small stone from a grave or threshold, warmed in the palm.

Choose one: ease a hard crossing (target gains +1 die to rolls involving change, decision, or emotional steadiness this scene) or find a hidden path (gain +2 dice to Navigation or Investigation when seeking unconventional routes).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Unburdened Farewell (6 XP)}

\textbf{Tags:} [HEAL], [CLEANSE], [DEBT], [FEAR]

\textbf{Threshold:} A personal letter never sent, burned in a copper dish.

Speak a final goodbye to someone not present—a lost lover, a dead parent, a friend you wronged. Clear 1 Fatigue and remove one psychological Condition (e.g., Haunted, Guilty) caused by that relationship. The GM gains 1 SB as the words drift into the threshold.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Passing Flock (8 XP)}

\textbf{Tags:} [GUIDANCE], [KNOWLEDGE], [HEAL], [AREA], [RISK]

\textbf{Threshold:} Black wool and a personal token of the one to be guided.

Choose one: open a brief passage through shadow-realms (target bypasses one insurmountable obstacle) or call shadow-sheep that carry fragments of weighty moments (retrieve lost details, soothe grief, or glean counsel from what was).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Steadfast Witness (7 XP)}

\textbf{Tags:} [WITNESS], [DEBT], [TRUTH], [THRESHOLD]

\textbf{Threshold:} A dying person's last words, spoken aloud and witnessed.

Record the dying's final testament. The words are binding on the living. Any who break them suffer -1 die to all actions for one week unless they atone. The GM gains 1 SB (the threshold remembers).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Unburdened Crossing (13 XP)}

\textbf{Tags:} [DEATH], [GUIDANCE], [FATE], [THRESHOLD]

\textbf{Threshold:} A dying person, a witness, and a grey wool cord.

Guide a willing creature across the threshold of death. They die peacefully, without fear. Before they pass, they may speak one final truth. That truth is considered witnessed; it cannot be denied by any who hear it. You gain a permanent memory of their last words, which may be used as a \textit{Truth Compass} (+1 die to any roll directly related to fulfilling their final wish).

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Comforting Presence (4 XP)}

\textbf{Requirement:} The Pale Shepherd

Once per scene, when you touch a creature suffering from Fear, Guilt, or Grief, you may remove that Condition. The creature clears 1 Fatigue if the Condition was particularly heavy (GM discretion).

\subsubsection{Final Witness (6 XP)}

\textbf{Requirement:} The Pale Shepherd, Spirit 3+

Once per session, when a sentient creature dies within Near range, you may ask the GM one question about that creature's final thoughts or unfulfilled wishes. The GM must answer truthfully.

\subsubsection{Threshold Keep (8 XP)}

\textbf{Requirement:} The Pale Shepherd, Spirit 3+

You may sense any active threshold (door, gate, bridge, border) within Far range, and you know its mood (welcoming, hostile, sorrowful, waiting). This is a passive ability.

\subsection{Roleplaying a Death-Warden}

\begin{itemize}
\item You do not fear death. You have seen it too many times. You respect it.
\item You may be called to witness deathbed testaments, to guide the dying, or to comfort the grieving. Your presence is a gift—and a reminder.
\item You may struggle with the weight of farewells. Every goodbye leaves a mark. The grey cloak grows heavier.
\item You may be asked to let someone go—a patient, a friend, yourself. The Shepherd does not hold. The Shepherd accompanies.
\end{itemize}

\subsection{Sample NPC: Sister Anja of the Grey Cloak}

\textbf{Archetype:} Death-Warden

\textbf{Patron:} The Pale Shepherd

\textbf{Codex:} Grey cloak, each fold a farewell witnessed

\textbf{Thiasos:} A grey bird that sings when someone is about to die

\textbf{Personality:} Calm, gentle, inexhaustibly sad. She has guided three hundred souls across the threshold. She remembers every name.

\textbf{Conflict:} A child is dying. The child's mother begs Anja to save her. Anja cannot. She can only accompany. The grey bird sings.

\subsection{GM Guidance: Intrusions for The Pale Shepherd}

\begin{itemize}
\item A ghost the Death-Warden once guided appears at an inopportune moment, asking for help with unfinished business.
\item The Death-Warden's grey cloak grows heavier. A soul is not passing; it is clinging. The threshold is blocked.
\item A funeral the Death-Warden attended is suddenly disputed—someone claims the deceased was not of sound mind when they spoke their final truth.
\item The Death-Warden dreams of a door they have never seen. On the other side, someone is waiting. The Shepherd does not say who.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
