\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{xcolor}

% Page Setup
\geometry{margin=1in}
\setlength{\headheight}{15pt}
\fancyhf{}
\rhead{\thepage}
\lhead{Fate's Edge: Ashes of the Infernal Accord}
\pagestyle{fancy}

% Section Styling
\titleformat{\section}{\large\bfseries\uppercase}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\itshape}{}{0em}{}

% Custom Commands
\newcommand{\patronbox}[1]{\noindent\textbf{#1}\\}
\newcommand{\clockbox}[2]{\noindent\textbf{#1 [#2]}\\}
\newcommand{\gmnote}[1]{\textit{GM Note: #1}}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Ashes of the Infernal Accord},
    pdfauthor={Fate's Edge GM},
}

\begin{document}

\begin{center}
    {\LARGE \textbf{Ashes of the Infernal Accord}}\\[0.5em]
    {\large An Epic Tier Adventure Module for Fate's Edge}\\[1em]
    {\itshape Tier: Epic (220+ XP) \quad Themes: Cosmic Bargain, Infernal Politics, Unreliable Alliances}
\end{center}

\section*{GM Quick Sheet: At the Table Flow}

\textbf{Hook:} Maelstraeus offers a deal to stop Mor'iraath's expansion. Accepting grants immediate benefit but \textbf{+4 Obligation} to Maelstraeus.

\textbf{Act I: Descent}
\begin{itemize}[leftmargin=*]
    \item Journey to City of Cinder (modified travel rules).
    \item \textbf{Early Encounter:} Minor Mor'iraath agent or Livaea's whisper test.
    \item \textbf{Stakes:} Preventing infernal influence spillover into mortal realm.
\end{itemize}

\textbf{Act II: The Infernal Bazaar}
\begin{itemize}[leftmargin=*]
    \item Use \textbf{City of Cinder Generator} for locations/encounters.
    \item \textbf{Social Core:} Navigate factions, gather info, make deals.
    \item \textbf{Livaea's Threads:} \textit{Foreshadow early} -- tempting offers that feel ``too good,'' dreams of perfect outcomes, NPCs oddly fixated on player desires.
    \item \textbf{Key Clocks:}
    \begin{itemize}
        \item \clockbox{Mor'iraath's Expansion}{8} (Stops his plan)
        \item \clockbox{Cinder Market Stability}{6} (Prevents city collapse)
        \item \clockbox{Livaea's Manipulation}{6} (Reveals her full scheme)
    \end{itemize}
\end{itemize}

\textbf{Act III: The Fractured Court}
\begin{itemize}[leftmargin=*]
    \item Choose approach: Direct Confrontation or Infernal Subterfuge.
    \item \textbf{Livaea's Reveal:} Mid-to-late act -- manipulation exposed.
    \item \textbf{Hard Choices:} Complete Maelstraeus' goal (risk his hidden agenda) or defy him (incur wrath).
\end{itemize}

\textbf{Resolution:} Success = Contain threat + rewards + lasting consequences. Failure = Major front advancement, persistent infernal influence.

\section{Patron Entry: Mor'iraath, the Demon of Fire and Destruction}

\patronbox{Lore:} Mor'iraath is not merely a lord of flame, but the \textit{principle} of entropic consumption through commerce and conflict. Where Maelstraeus builds webs of interdependence, Mor'iraath seeks to reduce all value to ash. He rules the City of Cinder not through law, but through the inevitability of burnout, the finality of a deal gone wrong, and the allure of power that consumes its wielder. He is the patron of those who would burn the world for profit or pleasure, the ender of cycles, and the devourer of bargains. His sigil is a contract in flames that rewrites itself with each reading.

\patronbox{Patron's Gift (Imbuement):} Once per scene as an action (cost: 1 Boon; requires Thiasos), touch an item to imbue it until scene end with +1 Melee and +1 Skirmish or Command (when leveraging intimidation/destruction) (Thematic). Push It: extend for one extra scene by marking +1 Obligation. (Same rules as core Invoker Gift).

\patronbox{Low Rites}
\begin{itemize}[leftmargin=*]
    \item \textbf{Rite of the Burning Ledger (Low):} Duration: Instant; Range: Near; Stacking: No.
    \begin{itemize}
        \item \textbf{Materials:} A scrap of parchment singed at the edges.
        \item \textbf{Effect:} Target suffers -1 die to one economic/social transaction roll (negotiation, trade, payment) this scene. Gain +1 die to an attack or destructive action this scene.
        \item \textbf{Invoke:} 1 action; mark +1 Obligation.
        \item \textbf{Push It:} The effect lingers subtly on the target's dealings for the rest of the session; mark 1 SB (Clubs) as financial records become... difficult.
    \end{itemize}
    \item \textbf{Rite of Ashen Sight (Low):} Duration: Scene; Range: Self; Stacking: No.
    \begin{itemize}
        \item \textbf{Materials:} Rub ash into the eyes (harmless but stings).
        \item \textbf{Effect:} Gain +1 die to Notice/Investigate actions to find weakness, fault lines, or instability in objects, structures, or social groups. Gain +1 Effect when targeting such weaknesses.
        \item \textbf{Invoke:} 1 action; mark +1 Obligation.
        \item \textbf{Push It:} Also see the ``price'' or ``cost'' associated with a target (in SB, Fatigue, or narrative terms); the sight is disturbing, mark 1 SB (Diamonds) as the knowledge unsettles you.
    \end{itemize}
\end{itemize}

\patronbox{Standard Rites}
\begin{itemize}[leftmargin=*]
    \item \textbf{Rite of the Conflagration Clause [AREA][FOLLOW-UP] (Standard):} Duration: Scene; Range: Zone; Stacking: No.
    \begin{itemize}
        \item \textbf{Materials:} A small fire set upon a contract or valuable item.
        \item \textbf{Effect:} Designate a zone. Enemies in the zone suffer -1 die to actions requiring patience, stability, or cooperation. Once per round, as an action, you may cause a small, damaging explosion (treat as a hazard, Cap 2-3, Harm 1-2) anywhere within the zone.
        \item \textbf{Invoke:} 1 action; mark +1 Obligation.
        \item \textbf{Push It:} Increase the explosion's Cap by +1 and Harm by +1; mark +1 Obligation.
    \end{itemize}
    \item \textbf{Rite of the Infernal Audit [REVEAL][CURSE] (Standard):} Duration: Scene; Range: Near; Stacking: No.
    \begin{itemize}
        \item \textbf{Materials:} A ledger opened to a random page, held aloft.
        \item \textbf{Effect:} Target must make a Resolve test (DV 3) or suffer -1 die to all actions due to stress/anxiety. On a failure, you learn one hidden financial/legal obligation or secret debt they are trying to hide. If they are currently benefiting from a deal or boon, it becomes [COMPROMISED] or suffers a minor setback (GM discretion).
        \item \textbf{Invoke:} 1 action; mark +1 Obligation.
        \item \textbf{Push It:} The target suffers Harm 1 (Stress) on a failed test; mark +1 Obligation.
    \end{itemize}
\end{itemize}

\patronbox{High Rites}
\begin{itemize}[leftmargin=*]
    \item \textbf{Rite of the Final Sale [BANISH][UNWARD] (High):} Duration: Instant; Range: Near; Stacking: No.
    \begin{itemize}
        \item \textbf{Materials:} A contract signed in the target's blood (or symbolic equivalent) and burned.
        \item \textbf{Effect:} Target a visible Outsider or significant magical effect/ward. Test DV = Cap or Tier. On Hit: add +DV segments to its Leash/Exit Tally or immediately suppress/unravel the effect; on Partial: add +1 segment or apply a minor suppression/setback; on Miss: the contract backfires, apply 1 SB (Spades) to the caster.
        \item \textbf{Invoke:} 1 action; mark +2 Obligation.
        \item \textbf{Push It:} If successful, the target/entity is not just banished/suppressed but \textit{consumed}, potentially gaining a benefit for the caster (like clearing 1 Obligation or gaining 1 Boon), but mark 2 SB (Diamonds) as the act draws the attention of other, potentially more powerful, entities or creates a lingering void/malignancy.
    \end{itemize}
    \item \textbf{Rite of the Worldfire Gambit [WARD][OMEN] (High - Epic Tier):} Duration: Extended (Significant Time); Range: Zone; Stacking: No.
    \begin{itemize}
        \item \textbf{Materials:} A massive bonfire fed by significant symbolic items (wealth, oaths, structures).
        \item \textbf{Effect:} Consecrate a vast zone (like a city district or battlefield). For the next session: All transactions, oaths, and value-based interactions within are inherently unstable and carry risk. Allies gain +1 Effect to destructive actions but suffer -1 die to cooperative or stabilizing actions. Enemies suffer the opposite. Start a ``Consumption Clock'' [8] (representing the area's value/energy being burned away).
        \item \textbf{Invoke:} Extended ritual; mark +3 Obligation.
        \item \textbf{Push It:} The effect is permanent (until countered) and the Consumption Clock ticks faster (+1/2 automatically per scene); mark +2 Obligation.
    \end{itemize}
\end{itemize}

\patronbox{Obligation:} Starts at 6 for Tier III characters, scaling appropriately for higher tiers.

\patronbox{Rivalry:} Maelstraeus (Obvious Antagonism - deals vs. destruction), Livaea (Subtle Manipulation - desire can feed or be consumed by fire).

\patronbox{Hard Obligation Triggers (Module-Specific):}
\begin{itemize}[leftmargin=*]
    \item \textbf{Obligation 8+ Segments:} A \textbf{Silent Collector} arrives to enforce an ``emergency clause'' in your pact -- demands an immediate, costly service or marks you as a debtor in default (\textbf{GM chooses: +2 Harm, lose asset, or mark permanent Condition}).
    \item \textbf{Obligation 10+ Segments:} \textbf{Maelstraeus calls in his debt immediately} -- you must perform a significant service for him \textit{right now}, potentially derailing your current objective or forcing a choice between patrons. Failure risks permanent enmity.
\end{itemize}

\newpage

\section{Region Generator: The City of Cinder - ``The Infernal Bazaar''}

\patronbox{Theme \& Atmosphere:} The City of Cinder is a sprawling, ever-shifting marketplace built from the calcified remnants of broken dreams and failed ambitions. It exists in the Realm of Fire, where the currency is not gold, but concepts: fragments of soul, moments of joy traded for fleeting power, oaths made in desperation, and the ashes of burnt-out lives. The architecture is a chaotic blend of grandiose auction houses, cramped pawn shops, and vast foundries where reality itself is bartered. The air thrums with the whispered terms of a million deals, and the ground is paved with the discarded contracts of the damned. Everything has a price, and everything can be bought... for a cost.

\patronbox{(Auction House/Forge/Alley):} The Grand Exchange, a vast dome where screams are the bidding currency; The Crucible Foundry, where memories are melted down into ingots of raw emotion; The Alley of Discarded Wishes, where the truly desperate come to trade their last hope.

\patronbox{Spades -- Places (categories)}
\begin{multicols}{2}
\begin{itemize}[leftmargin=*]
    \item 2. (Despair/Bargain) The Alley of Discarded Wishes
    \item 3. (Commerce/Fire) The Crucible Foundry
    \item 4. (Deception/Heat) The Grand Exchange
    \item 5. (Destruction/Value) The Vault of Broken Pacts
    \item 6. (Commerce/Pressure) The Loan Quarter
    \item 7. (Fire/Competition) The Arena of Wills
    \item 8. (Destruction/Luxury) The Palace of Ashes
    \item 9. (Commerce/Corruption) The Black Market Bourse
    \item 10. (Fire/Power) The Heartfire Conduit
    \item J. (Despair/Intrigue) The Archive of Regrets
    \item Q. (Commerce/Fear) The Fearmonger's Row
    \item K. (Destruction/Inevitability) The End Zone
    \item A. (Fire/Consumption) The Maw
\end{itemize}
\end{multicols}

\patronbox{Hearts -- People \& Factions (categories)}
\begin{multicols}{2}
\begin{itemize}[leftmargin=*]
    \item 2. (Desperation/Bargain) Pawnbroker of Last Resorts
    \item 3. (Commerce/Fire) Ember-Scribe
    \item 4. (Deception/Value) The Counterfeiter of Souls
    \item 5. (Destruction/Competition) Pit-Fighter
    \item 6. (Commerce/Debt) The Silent Collector
    \item 7. (Fire/Luxury) The Ash Merchant
    \item 8. (Despair/Corruption) The Whisperer of Better Deals
    \item 9. (Commerce/Fear) The Tax Assessor
    \item 10. (Destruction/Power) The Forge-Master
    \item J. (Desperation/Intrigue) The Regret-Seller
    \item Q. (Commerce/Fear) The Fear-Eater
    \item K. (Fire/Inevitability) The Final Arbiter
    \item A. (Destruction/Consumption) Mor'iraath's Avatar
\end{itemize}
\end{multicols}

\patronbox{Clubs -- Complications/Threats (categories)}
\begin{multicols}{2}
\begin{itemize}[leftmargin=*]
    \item 2. (Despair/Bargain) Contract Snare
    \item 3. (Commerce/Fire) Market Crash
    \item 4. (Deception/Heat) Price Surge
    \item 5. (Destruction/Value) Asset Impoundment
    \item 6. (Commerce/Pressure) Credit Collapse
    \item 7. (Fire/Competition) Hostile Takeover
    \item 8. (Destruction/Luxury) Luxury Plague
    \item 9. (Commerce/Corruption) Bribe Demand
    \item 10. (Despair/Fear) Panic Sell-Off
    \item J. (Commerce/Intrigue) Market Manipulation
    \item Q. (Fire/Fear) Inflationary Spiral
    \item K. (Destruction/Inevitability) Debt Avalanche
    \item A. (Fire/Consumption) The Maw Awakens
\end{itemize}
\end{multicols}

\patronbox{Diamonds -- Rewards/Leverage (categories)}
\begin{multicols}{2}
\begin{itemize}[leftmargin=*]
    \item 2. (Despair/Bargain) Debt Forgiveness Voucher
    \item 3. (Commerce/Fire) Ember-Infused Alloy
    \item 4. (Deception/Value) Counterfeit Charter
    \item 5. (Destruction/Competition) Pit-Fighter's Bond
    \item 6. (Commerce/Debt) Favor Token
    \item 7. (Fire/Luxury) Ash-Wrought Relic
    \item 8. (Despair/Corruption) Soul Shard
    \item 9. (Commerce/Fear) Fearless Mantle
    \item 10. (Destruction/Power) Crucible Key
    \item J. (Desperation/Intrigue) Regret Nullifier
    \item Q. (Commerce/Fear) Market Stabilizer
    \item K. (Fire/Inevitability) Contract of Dominion
    \item A. (Destruction/Consumption) Maw's Blessing
\end{itemize}
\end{multicols}

\patronbox{Quick use notes}
\begin{itemize}[leftmargin=*]
    \item Draw until you have all four suits: Spade = place, Heart = actor, Club = pressure, Diamond = leverage. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
    \item Diamonds are codified outcomes (rewards/leverages) that change position rather than call for a roll.
    \item If any A appears, echo fire \& consumption motifs---contracts that rewrite themselves, items that burn without fuel, value that vanishes upon touch.
\end{itemize}

\patronbox{Additional Features}
\begin{itemize}[leftmargin=*]
    \item \textbf{Value Flux:} The \textit{value} of things (items, information, even people) fluctuates wildly and often illogically. A common pebble might fetch a fortune one moment, be worthless the next. Social Position and Effect in Cinder are heavily influenced by what the market currently \textit{desires}.
    \item \textbf{The Ember-Scribe's Eye:} Every deal made in Cinder is recorded. Secrets are commodities. Knowing the right question to ask the Ember-Scribe (Wits + Lore, DV 4, generates SB) can reveal crucial information, but the knowledge often comes with a price or obligation.
    \item \textbf{Soulfire:} The primary energy source. It burns bright but consumes the user's essence over time. Prolonged presence in Cinder or use of Cinderite magic/assets should tick a ``Soulfire Exposure'' clock for characters, leading to Fatigue, Conditions, or worse.
\end{itemize}

\patronbox{Thematic SB Spend Table (Cinder-Specific)}
\textbf{Minor Complications (1 SB)}
\begin{itemize}[leftmargin=*]
    \item Contractual Snag: A deal you made has a hidden clause revealed at the worst moment.
    \item Price Spike: The cost of a needed item/service suddenly doubles.
    \item Market Whispers: Rumors spread about you, complicating social interactions.
    \item Ember Sickness: Brief Fatigue due to exposure to raw infernal energy.
    \item Scapegoat Demand: Authorities or a mob needs someone to blame for a recent failure; you fit the description.
\end{itemize}

\textbf{Moderate Setbacks (2 SB)}
\begin{itemize}[leftmargin=*]
    \item Asset Seizure: A valuable item or asset is impounded or claimed by a creditor.
    \item Debt Collector: A Silent Collector or similar arrives to enforce an old or unexpected obligation.
    \item Market Panic: A localized crash or surge affects your plans.
    \item Rival Sabotage: A competing faction or individual actively works against you.
    \item Identity Theft: Your name/credit is used fraudulently; clearing it is difficult.
\end{itemize}

\textbf{Serious Trouble (3 SB)}
\begin{itemize}[leftmargin=*]
    \item Contractual Trap: You are bound by a deal you didn't fully understand or was misrepresented.
    \item Soulfire Burn: Suffer Harm 1 (Burn) from overexposure or backlash.
    \item Major Rival Ascendant: A powerful Cinderite faction gains significant ground, threatening your position/goals.
    \item Patron's Wrath (Mor'iraath): A direct sign of the Demon's displeasure; immediate complication tied to fire/destruction.
    \item Clock Tick: Advance a relevant campaign/front clock by 1 segment.
\end{itemize}

\textbf{Major Turns (4+ SB)}
\begin{itemize}[leftmargin=*]
    \item Market Collapse: A significant sector of the City of Cinder fails, causing widespread chaos.
    \item The Maw Hungers: A district is consumed by the central void, forcing mass evacuation or adaptation.
    \item Avatar of Destruction: Mor'iraath's Avatar or a major champion arrives to deal with you personally.
    \item Livaea's Web Tightens: A major manipulation by Livaea is revealed, changing the fundamental nature of the conflict.
    \item Soulfire Plague: Raw, uncontrolled energy runs rampant, affecting everyone and everything unpredictably.
\end{itemize}

\newpage

\section{Faction Deal Clocks (Dynamic Politics)}

\patronbox{1. The Ember Merchants' Guild Influence [6]}
\begin{itemize}[leftmargin=*]
    \item \textbf{Tick +1/2:} A successful deal, a new luxury trend, effective lobbying.
    \item \textbf{Tick -1/2:} A scandal, product recall, loss of key member.
    \item \textbf{Fills (Positive):} Gain +1 die on transactions involving rare/soulfire materials. Their protection makes one Asset or Follower Neglected/Compromised status clear.
    \item \textbf{Fills (Negative):} They corner the market -- prices spike, access denied. Start a ``Guild Monopoly Pressure [4]'' clock affecting the party's operations.
\end{itemize}

\patronbox{2. The Arena of Wills Champions' Favor [6]}
\begin{itemize}[leftmargin=*]
    \item \textbf{Tick +1/2:} A champion wins a major bout, a gladiatorial spectacle boosts morale.
    \item \textbf{Tick -1/2:} A popular champion is defeated/disgraced, internal disputes.
    \item \textbf{Fills (Positive):} Gain a temporary Cap 3 Pit-Fighter follower for one major conflict. Allies gain +1 die to Command rolls involving strength/fear.
    \item \textbf{Fills (Negative):} A champion goes rogue/mad, causing havoc. Start an ``Arena Riot [6]'' clock threatening a district.
\end{itemize}

\patronbox{3. The Archive of Regrets' Knowledge Trade [6]}
\begin{itemize}[leftmargin=*]
    \item \textbf{Tick +1/2:} A valuable secret is traded, a historical mystery solved.
    \item \textbf{Tick -1/2:} A forbidden secret is leaked, causing panic or attracting unwanted attention.
    \item \textbf{Fills (Positive):} Gain one powerful piece of forbidden lore (like a major ritual secret or a vulnerability of a key antagonist) usable once. Clear 1 segment from any Investigation-based clock.
    \item \textbf{Fills (Negative):} A catastrophic truth is revealed (perhaps about a PC's past or a Patron's true goal), causing immediate Harm 2 (Stress) to one character and starting a ``Revelation Fallout [4]'' clock.
\end{itemize}

\section{Livaea's Foreshadowing Integration}

Throughout Acts I and II, weave Livaea's influence subtly:
\begin{itemize}[leftmargin=*]
    \item \textbf{Early Deals:} An offer appears that perfectly solves a short-term problem but has vague long-term clauses. If taken, mark 1 SB (Hearts) -- the ``favor'' will be called in later under bad circumstances.
    \item \textbf{Dreams/Sights:} Characters might catch glimpses of their deepest desires fulfilled \textit{if they just make one more deal}. These visions are compelling but always show a dark edge upon closer inspection (e.g., ruling a kingdom, but it's built on ash and screams).
    \item \textbf{NPC Behavior:} A seemingly helpful Cinderite NPC becomes oddly fixated on one PC's personal goals or flaws, offering advice that steers them toward Mor'iraath's sphere of influence under the guise of helping.
    \item \textbf{Market Whispers:} Rumors circulate about a ``Patron of Perfect Outcomes'' or ``She Who Grants True Desires'' -- planting the seed that there's another player beyond Maelstraeus and Mor'iraath.
\end{itemize}
This ensures her reveal in Act III feels like the uncovering of a masterfully laid plan rather than a last-minute twist.

\end{document}
