\section{The Banner Swap in the Wind}
\emph{A Black Banners Adventure for Fate's Edge}

\subsection*{Adventure Overview}
\begin{tabular}{@{}ll@{}}
\textbf{Tier:} & II–III \\
\textbf{Length:} & 2–3 sessions \\
\textbf{Theme:} & Banner politics, village defense, condotta manipulation \\
\textbf{Hook:} & A Vilikari village caught between two rival banner companies offers the party a chance \\
& to play the companies against each other for the village's survival.
\end{tabular}

\subsection*{Setup: The Village of Saltmere}
Saltmere sits in a strategic valley where two trade routes converge. The \textbf{Iron Lanterns} (siege specialists who promise ``no fire in hospitals'') and the \textbf{Thornwrights} (sappers/road-cutters who price bridges by the breath) hold \emph{conflicting charters} to secure the valley. Village elders approach the PCs to manipulate both banners into terms that spare Saltmere.

\paragraph{Saltmere at a Glance}
\begin{itemize}[leftmargin=*]
  \item \textbf{Assets:} Winter granaries, shrine-hospital (old stone), river ford, timberyards, waymarket green.
  \item \textbf{Strings:} Ledger of Small Kindnesses (debts owed), ford-right witness-stone, bell-token set (for parley).
  \item \textbf{Clocks:} \emph{Granary Seizure [4]}, \emph{Shrine Fire Rumors [4]}, \emph{Village Panic [6]}.
\end{itemize}

\subsection*{Key NPCs}
\paragraph{Village Leadership}
\textbf{Matron Kessa} — Vilikari elder versed in banner law; speaks in ledgers and exits.\\
\textbf{Trader Nilo} — Factor who keeps the Ledger of Small Kindnesses; quiet leverage artist.\\
\textbf{Young Vek} — Scout who tracks troop movements; eager but careful.

\paragraph{Banner Companies}
\subsubsection*{The Iron Lanterns (Siege Specialists)}
\textbf{Captain Thorne Ironwright} — Methodical, honor-bound, obsessed with ``clean sieges.''\\
\textbf{Quartermaster Bray} — Counts twice; hates waste; knows who is hungry.\\
\textbf{Strings:} Bridge priority stamps; siege contracts; hospital sanctuary rights.\\
\textbf{Company Tracks (scene-facing):} Payday [\;], Supply [\;], Cohesion [\;].
\medskip

\subsubsection*{The Thornwrights (Sappers / Road-Cutters)}
\textbf{Captain ``Cut'' Marrick} — Reckless innovator who prices everything in breaths.\\
\textbf{Engineer Tessa Chain} — Structural genius with a gambling problem.\\
\textbf{Strings:} Road contracts; bridge toll ledgers; explosive licenses.\\
\textbf{Company Tracks (scene-facing):} Payday [\;], Supply [\;], Cohesion [\;].

\subsection*{Session 1: The Two Charters}
\paragraph{Opening Scene: Village Council}
When the PCs arrive, the Lanterns have a ``peaceful blockade'' on the north road; the Thornwrights dig ``observation trenches'' on the south ridge. Both claim to protect the village—on paper. The winter grain sits between them; neither allows unescorted movement.

\paragraph{Investigation Beats (pick any order)}
\begin{enumerate}[leftmargin=*]
  \item \textbf{Charter Analysis} (\emph{Wits+Lore or Diplomacy (Broker) DV 3}): Overlapping claims with \emph{different end terms}: Lanterns ``until \emph{first frost}''; Thornwrights ``until \emph{the bridge is paid}.'' Note the \emph{witnesses} differ.
  \item \textbf{Ledger of Small Kindnesses} (\emph{Wits+Investigation DV 4}): Kessa/Nilo reveal prior debts: Lanterns once used the shrine as triage (owe sanctuary upkeep); Thornwrights crossed free during spring thaw (owe toll).
  \item \textbf{Scout Reports} (\emph{Wits+Survey DV 3}): Vek notes Lanterns assemble siege mantlets ``just in case''; Thornwrights quietly mine the old bridge footings.
\end{enumerate}

\paragraph{The Opportunity}
Under camp Articles, \emph{prior debts} and \emph{sanctuary rights} must be honored. Kessa proposes the party present each company with evidence the other is breaching terms—pressing them toward withdrawal or a new charter.

\subsection*{Session 2: Playing the Companies}
\paragraph{Condotta Flip (Allegiances-in-Flux, framed by PC play)}
Roll 1d6 at scene starts or when PCs spring a gambit; interpret via their actions:
\begin{enumerate}[leftmargin=*,label=\Alph*.]
  \item \textbf{Charter Challenge:} Present evidence Thornwrights cache explosives in the shrine (\emph{Deception+Lore DV 4–5} or true if PCs confirm). Lanterns must defend sanctuary.
  \item \textbf{Debt Reminder:} Confront Thornwrights with bridge toll debt (\emph{Diplomacy+Broker DV 3}); shifts their Position –1 until redressed.
  \item \textbf{False Orders:} Plant misrouting slips (\emph{Subterfuge+Deception DV 4}); start \emph{False Orders [4]} for that company.
  \item \textbf{Broker's Market:} Offer exclusive rights to alternate routes (\emph{Broker DV 3–4}); on hit, company accepts narrower claim.
  \item \textbf{Village Leverage:} Threaten to provision the rival (\emph{Presence+Command DV 3}); on partial, \emph{Village Panic +1}.
  \item \textbf{Article Violation:} Accuse of breaking Articles (e.g., lantern oath, toll oath) under witness (\emph{Presence+Sway DV 4}); on hit, force a Parley.
\end{enumerate}

\paragraph{Key Scenes (structure and stakes)}
\subsubsection*{Bell Dawn Parley} 
\emph{Presence+Sway DV 4 under witness; ring three bells to invoke sanctuary.}\\
Both captains must attend. PCs set speaking order (Position +1 if they present Ledger first). On hit, choose 1:
\begin{itemize}[leftmargin=*]
  \item Impose \emph{No-Seizure of Granaries [4]} as a shared Article (both companies gain DV +1 to resist looting SB).
  \item Extract agreement to \emph{Single Banner Hour} (only one company shows colors in town per day).
\end{itemize}
On miss, Keeper may spend SB: \emph{Authority Intervention} (local lord's reeve arrives) or \emph{Camp-Fever} (morale sag).

\subsubsection*{The False Ledger}
\emph{Wits+Deception DV 5 (or Wits+Lore DV 4 if based on real entries).}\\
Forge a \emph{patron addendum} implying a side-switch. On hit, target company \emph{Withdraws 1 ring} or loses \emph{Cohesion +1}. On partial, it merely \emph{Hesitates}: Position –1 next scene.

\subsubsection*{The Village Exchange}
\emph{Broker or Diplomacy DV 3–4, contested by watching officers.}\\
Run a public market where each company bids for exclusive rights. On hit, the village secures \emph{Bridge Peace [4]} (ticks each day of calm). On miss, \emph{Granary Seizure +1}.

\subsection*{Session 3: Resolution \& Consequences}
\paragraph{The Banner Swap Crisis}
Roll Allegiances-in-Flux once mid-negotiation; apply:
\begin{itemize}[leftmargin=*]
  \item \textbf{1–2 (Backlash):} Both companies realize manipulation; \emph{False Orders [4]} becomes \emph{Skirmish Spiral [6]} unless PCs de-escalate (DV 4).
  \item \textbf{3–4 (Bidding War):} One company escalates terms (dangerous exclusives); \emph{Payday +1} for the other as soldiers grumble.
  \item \textbf{5–6 (Third Claimant):} A local lord/merchant house presents a writ; start \emph{Political Pressure [4]}.
\end{itemize}

\paragraph{Climax Options (choose or mix)}
\subsubsection*{1. Charter Synthesis}
\emph{Broker+Petition DV 5 (teamwork).}\\
Draft a new shared charter splitting toll and ford rights by season/hours. On hit, both patrons accept (gain \emph{Broker's Favor} x2). On partial, one company accepts; the other demands \emph{Indemnity [4]}.

\subsubsection*{2. The Village Gambit}
\emph{Lore+Diplomacy DV 4 under witness.}\\
Invoke \emph{Vilikari Ledger Law}: old debts void present claims. On hit, both charters \emph{narrow} to outer rings; village keeps core. On miss, \emph{Authority Intervention}—the reeve freezes movement (Supply +1).

\subsubsection*{3. The Third Way}
\emph{Command+Presence DV 4.}\\
Propose joint defense versus bandits/raiders/storm. If accepted, start \emph{Shared Defense [6]}; when filled, both companies leave honor-intact, village gains \emph{Watch Rights}.

\subsection*{Rewards \& Consequences}
\paragraph{Success Outcomes}
\begin{itemize}[leftmargin=*]
  \item \textbf{Clean Victory:} Both companies withdraw with honor; Saltmere safe. PCs gain \emph{Broker’s Favor} (Lanterns \& Thornwrights).
  \item \textbf{Pragmatic Victory:} One company leaves; the other remains as protector on narrowed charter. PCs gain a \emph{Village String} (free ford use; Position +1 at Saltmere).
  \item \textbf{Innovative Victory:} New charter adopted; PCs become known as \emph{banner mediators} (Reputation tag).
\end{itemize}

\paragraph{Failure Consequences}
\begin{itemize}[leftmargin=*]
  \item \textbf{Village Lost:} Saltmere burned or emptied; PCs gain a \emph{Black Mark} with brokers.
  \item \textbf{Banner War:} Companies clash; choose a side or flee. Start \emph{Skirmish Spiral [6]}.
  \item \textbf{Patron Wrath:} Offended patrons pursue reprisals; \emph{Political Pressure [4]} follows the party.
\end{itemize}

\subsection*{Mechanical Hooks}
\paragraph{Key Dice Rolls}
\begin{itemize}[leftmargin=*]
  \item \textbf{Charter Reading:} Wits+Lore (DV 3–4) to spot conflicting clauses/witness gaps.
  \item \textbf{Banner Law:} Presence+Sway (DV 4) under witness/oaths; Position +1 if bells rung.
  \item \textbf{Village Negotiation:} Diplomacy or Broker (DV 3–4) with competing interests present.
  \item \textbf{False Evidence:} Deception+Lore (DV 4–5) to forge convincing addenda.
  \item \textbf{Supply Cut/Restore:} Tactics+Streetwise or Engineering (DV 3–4) to hinder/help lines.
\end{itemize}

\paragraph{SB Menu (Keeper)}
\begin{itemize}[leftmargin=*]
  \item \textbf{False Orders} (colors cross at the ford)
  \item \textbf{Ally Turns} (officer flips when shamed)
  \item \textbf{Cut Supply Lines} (granary cordon tightens)
  \item \textbf{Camp-Fever} (restless troops, morale dips)
  \item \textbf{Gear Failure} (siege mantlet collapse; mine misfire)
  \item \textbf{Authority Intervention} (lord’s reeve with writs)
  \item \textbf{Winter Omens} (first frost early; bridges ice)
\end{itemize}

\subsection*{Adventure Seeds for Continuation}
\begin{enumerate}[leftmargin=*]
  \item \textbf{The Broker’s Debt:} The scribe who drafted the conflicting charters hires the PCs to unwind a bigger snarl.
  \item \textbf{Village Ascendancy:} Saltmere’s stance inspires a league of valley villages; negotiate a \emph{Blue Moot}.
  \item \textbf{Patron Intrigue:} The clash masked treasury games; follow the writs upstream.
  \item \textbf{Banner Schism:} Junior officers question the Articles; a reform faction courts the PCs.
\end{enumerate}

\subsection*{GM Notes}
\paragraph{Tone \& Pacing}
Keep focus on clever manipulation over direct battle. End each session with a \emph{clock visibly advancing}: discovery \(\rightarrow\) leverage \(\rightarrow\) crisis.

\paragraph{Table Tools}
\begin{itemize}[leftmargin=*]
  \item Keep \emph{Payday}, \emph{Supply}, \emph{Cohesion} for both companies visible (index cards or sidebar).
  \item Place a handbell on the table; when rung, sanctuary applies and social Position bumps for oathful conduct.
  \item Use \emph{Bowl then Board}: first fairness (who is harmed), then geometry (who holds which lane).
\end{itemize}

\subsection*{Handouts \& Props (printables)}
\paragraph{Conflicting Clauses (excerpt)}
\begin{quote}\small
``The Company of the Iron Lanterns shall observe the Sanctuary of the Saltmere Shrine-House; no fire nor powder shall be borne within twenty paces, and no man under arms shall pass her lintel save to bear the wounded. This charge endures until the \textit{first frost} under witness of the bell.''\\
``The Thornwright Company shall hold the ford and bridge-approaches for \textit{the price of the bridge}, to be reckoned in breaths at the gate, and may lay any engine or charge needful to assure the price, so long as no force impede the due commerce of the road.''
\end{quote}

\paragraph{Bell Dawn Rite (card text)}
\begin{quote}\small
Ring thrice at dawn. All banners furl in the green. Parley under witness: no seizure, no shot, no shout. Speak plain; record both said and meant.
\end{quote}

\subsection*{Quick Reference: Scene DVs}
\begin{center}\small
\begin{tabular}{lll}
\toprule
\textbf{Scene} & \textbf{Primary Test} & \textbf{DV / Notes} \\
\midrule
Charter Reading & Wits+Lore & 3--4; +1 die if Ledger present \\
Bell Dawn Parley & Presence+Sway & 4; Position +1 if bells/witness \\
False Ledger & Wits+Deception & 5 (4 if based on truth) \\
Village Exchange & Broker/Diplomacy & 3--4; contested by officers \\
De-escalate Spiral & Command+Tactics & 4; clear 1 from Skirmish Spiral \\
\bottomrule
\end{tabular}
\end{center}