\documentclass[11pt]{article}

% --- Packages ---
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=black,citecolor=black}
\setlist[itemize]{leftmargin=1.4em}
\setlist[enumerate]{leftmargin=1.6em}
\newcommand{\clock}[1]{\textsc{[#1]}}
\newcommand{\suit}[1]{\textit{#1}}

% --- Title ---
\title{\vspace{-1.5em}\textbf{Between the Knot and the Gate}\\
\large An Adventure Frame for \textit{Fate's Edge}}
\author{Tier II--III \quad Themes: Bargains vs.\ Mercy, Oaths, Thresholds, Memory}
\date{}

\begin{document}
\maketitle
\vspace{-1.25em}

\noindent\textbf{Primary Patrons:} Morag the Hag, The Pale Shepherd.\\
\textbf{Pitch:} A borderland duchy's pilgrimage gate---once guiding the dead to rest---is failing. Restless souls linger. Morag offers order through cruel bargains; the Pale Shepherd urges restoration and right endings. The party must decide what replaces the old order.

\section*{GM Quick Sheet (At-the-Table Flow)}
\textbf{Hook.} A noble heir returns memoryless from a failed funeral procession; hauntings begin.\\
\textbf{Immediate Offers.}
\begin{itemize}
  \item \textbf{Morag's Envoy:} Bind the duchy with thorned bargains for fast stability (deep cost).
  \item \textbf{Shepherd's Acolyte:} Restore the threshold gate through vigil, memory, and covenant (slow, compassionate).
\end{itemize}
\textbf{Global Clocks.}
\begin{itemize}
  \item \textbf{Stalled Passing} \clock{6} (dead linger; hauntings escalate).
  \item \textbf{Bargain Net Tightens} \clock{6} (oaths spread across civic life).
  \item \textbf{Gate Repair or Replacement} \clock{8} (relics, name, covenant).
\end{itemize}
\textbf{Hard Triggers.}
\begin{itemize}
  \item \textbf{Morag Obligation 10+:} Demand betrayal/sacrifice; refusal invites hexed interference.
  \item \textbf{Shepherd Obligation 9+:} Task to shepherd a specific lost soul; stalling worsens social Position for related scenes.
\end{itemize}

\section*{Act I --- Frayed Thresholds}
\textbf{Goals:} Learn causes, choose initial allegiance(s), expose stakes.
\begin{itemize}
  \item \textbf{Broken Procession.} Escort a funeral that fails at the old gate; lost souls manifest.\\
  Tests: Wits+Insight (oath residue), Spirit+Rites (soothe), Deception/Command (calm crowds).
  \item \textbf{Two Offers, One Cost.} Morag's agent pitches a \textit{Rite of the Thorned Bargain} for peace; the Shepherd's acolyte urges a \textit{Rite of the Memory Corral} to learn why the gate fails.
  \item \textbf{Crossroads Lead.} Use \textit{Rite of the Crossroads Whisper} to pick a path: investigate the \emph{Oath-Ledgers} (Morag) or the \emph{Buried Doorframe} (Shepherd).
\end{itemize}
\textbf{Starter Clocks.}
\begin{itemize}
  \item \textbf{Public Panic} \clock{4} (feeds other clocks).
  \item \textbf{Hidden Oath Cartel} \clock{6} (guild + magistrates).
  \item \textbf{Gate's Old Name Remembered} \clock{4} (Shepherd-leaning discovery).
\end{itemize}

\section*{Act II --- The Knot Tightens, The Gate Groans}
\textbf{Goals:} Unravel causes; choose methods of resolution. \emph{Run 2--3 threads.}
\begin{itemize}
  \item \textbf{The Oath Cartel.} Syndicate monetizes funerals via \textit{Rite of the Twisted Knot}.\\
  Scenes: ledger heist; \textit{Bone-Truth} interrogation; courtroom challenge under thorned bargain.
  \item \textbf{The Stolen Doorframe.} Relic from the first gate sits as décor, steeped in grief.\\
  Scenes: social infiltration (\textit{Borrowed Face}); ballroom \textit{Memory Corral}; \textit{Threshold Covenant} with the family.
  \item \textbf{The Lost Heir's Memory.} Missing memory holds the true keeper's name.\\
  Scenes: dream vigil (\textit{Gentle Passing} on the dying tutor); \textit{Shadow Flock} crypt traversal; Resolve to accept painful truth.
\end{itemize}
\textbf{Mid-Act Set Pieces.}
\begin{itemize}
  \item \textbf{Market of Masks.} Morag's fair-night; cheap bargains solve symptoms (temptation).
  \item \textbf{Vigil of Stillness.} Shepherd rite stabilizes a district for one night; costs Fatigue/Obligation.
\end{itemize}
\textbf{Evolving Clocks.}
\begin{itemize}
  \item \textbf{Bargain Net Tightens} \clock{6$\to$8} if bargains are used for civic order.
  \item \textbf{Gate Repair or Replacement} \clock{8} ticks with relics, name, and covenant.
\end{itemize}
\textbf{Reveals.} The keeper line is broken; the Cartel plugged the gap with profit-driven oaths. The relics (burned doorframe, water blessed by loss, the keeper's name) can re-bind the threshold.

\section*{Act III --- Binding or Passing}
\textbf{Goals:} Choose a lasting settlement for the duchy.
\begin{itemize}
  \item \textbf{Bind the City (Morag).} Enact a mass \textit{Thorned Bargain}: safety for obedience.\\
  Boss-as-System: Institutional bureaucracy (oath office) + Social movement (pro/anti-bargain).
  \item \textbf{Restore the Gate (Shepherd).} Perform \textit{Threshold Covenant} with a rightful keeper and community.\\
  Requires relics, public vigil, and accepting the cost (harm/condition/community burden).
  \item \textbf{Hybrid Third Way.} Limited time-boxed bargains to protect the weak, with sunset clauses after restoration; public oversight.
\end{itemize}
\textbf{Final Confrontations (Non-Combat Bosses).}
\begin{itemize}
  \item \textbf{Court of Oaths.} Debate, precedent, and ritual law determine civic order.
  \item \textbf{Funeral March.} Escort hundreds through the threshold; manage hazards, panic, predatory deals.
\end{itemize}
\textbf{Failure States.}
\begin{itemize}
  \item Gate collapses: \textbf{Stalled Passing} \clock{6$\to$10}; duchy becomes a liminal haunt.
  \item Oath-martial law: Bargain Net locks in; dissenters cursed or exiled.
\end{itemize}

\section*{Rewards \& Consequences}
\textbf{Morag Path.} Access to \textit{Cauldron's Boil} components; political leverage. Costs: Obligation spikes; risk \textbf{Bargain-Sight}; trust erosion.\\
\textbf{Shepherd Path.} One-use blessing akin to \textit{Shepherd's Mercy}; ``Good Death'' favors; safer funerary narratives. Costs: community burden, \textbf{Melancholy} risk, personal sacrifice clock.\\
\textbf{Hybrid.} Civic legitimacy; partial favor from both patrons; recurring auditors (Silent Collector / Acolyte).

\section*{Location Seeds}
\begin{itemize}
  \item \textbf{The Splintered Gate.} Ruined arch; candle-flames gutter upward; names whisper.
  \item \textbf{The Ledger Hall.} Thorn-carved lecterns; quills drink blood; oaths echo.
  \item \textbf{The Wakehouse.} Woven veils; memories condense like dew on glass.
  \item \textbf{The Manor of Borrowed Faces.} Portrait gallery that watches back.
  \item \textbf{The Crossroads Well.} Mouth-stones for whispers; offerings tied with knots.
\end{itemize}

\section*{NPCs (Quick Roles)}
\begin{itemize}
  \item \textbf{Archivist Avena (Shepherd-leaning).} Seeks honest endings; knows the old name.
  \item \textbf{Magistrate Thorn (Morag-leaning).} Efficient tyrant by paperwork; believes bargains are mercy.
  \item \textbf{Heir Lysa (Center).} Missing memory; potential threshold keeper.
  \item \textbf{Kettle-Mae (Morag's Envoy).} Charming; cruel when crossed.
  \item \textbf{Brother Tellen (Shepherd's Acolyte).} Exhausted; stubbornly kind.
\end{itemize}

\section*{Clocks \& Dials (Suggested Defaults)}
\begin{itemize}
  \item \textbf{Stalled Passing} \clock{6} (advances on public fear, failed vigils).
  \item \textbf{Bargain Net Tightens} \clock{6}/\clock{8} (advances when bargains maintain order).
  \item \textbf{Gate Repair or Replacement} \clock{8} (ticks with relics, name, covenant).
  \item \textbf{Public Panic} \clock{4} (feeds others).
  \item \textbf{Hidden Oath Cartel} \clock{6} (expose or dismantle).
  \item \textbf{Vigil of Stillness} \clock{4} (hold the line during rites).
\end{itemize}

\section*{SB Spend Menu (Thematic)}
\textbf{1 SB.} Contract snag; a mourner names a taboo; helpful memory shard stings.\\
\textbf{2 SB.} Oath loophole closes; vigil interrupted by a familiar face (\textit{Borrowed Face}); lost soul fixates on a PC.\\
\textbf{3 SB.} Spiteful curse (Harm 1--2); riot at the wake; false keeper declared.\\
\textbf{4+ SB.} Mass panic; oath purge; threshold backlash (zone-wide Stress/Fatigue).

\section*{Obligation Triggers (Foreground)}
\textbf{Morag 10+.} Demand betrayal/sacrifice; refusal invites manipulation and curses.\\
\textbf{Morag 12+.} Promises misfire; transformations backfire; pariah risk until a proving-quest.\\
\textbf{Shepherd 9+.} Shepherd a specific soul/memory; delaying worsens social Position.\\
\textbf{Shepherd 11+.} Threshold flicker: doors mislead to the Ways Between until penitent quest.

\section*{Resolution Epilogues (Pick a Tone)}
\textbf{Order at a Price.} Markets calm; funerals function; love grows cold.\\
\textbf{Mercy with Scars.} Hauntings end; grief honored; quiet burdens borne.\\
\textbf{Precarious Balance.} Civic oaths sunset; the gate tends itself---until next winter.

\vfill
\noindent\footnotesize\textit{Running Note:} Map Position to DV plainly (Controlled=DV2, Risky=DV3, Desperate=DV4--5+). Favor Position shifts over raw dice creep. Name costs before rolls. Keep total active clocks $\leq$~5 to avoid overload.

\appendix

\clearpage
\section{Patron Entry: The Pale Shepherd}
\label{patron:pale-shepherd}

\subsection*{Lore}
The Pale Shepherd is the guardian of thresholds, the watcher at the crossroads between life and death, sleep and waking, the mortal world and the Ways Between. Neither wholly benevolent nor malevolent, the Shepherd tends to the lost, the dying, and those who have wandered too far from their destined path. They appear as a figure in pale, weathered robes, often accompanied by a flock of shadow-sheep whose wool holds fragments of forgotten dreams and lost memories.  

The Shepherd speaks rarely, and when they do, their words are riddles that seem to answer questions not yet asked. They are invoked by those who seek guidance through liminal spaces, protection for the dying, or the recovery of lost things---though the Shepherd's help often comes with the price of a memory or a piece of one's future. Their sigil is a white ram's skull wreathed in mist.  

Those who serve the Pale Shepherd learn to navigate the spaces between states, to find paths where none seem to exist, and to shepherd others through their own dark nights. However, prolonged service renders the faithful melancholic, as they become attuned to the sorrow and loss that permeate all transitions.

\subsection*{Patron's Gift (Imbuement)}
Once per scene as an action (cost: 1 Boon; requires Thiasos), touch an item to imbue it until scene end with +1 Melee and +1 Stealth or Insight (when leveraging liminal spaces/transition moments) (Thematic). \emph{Push It:} extend for one extra scene by marking +1 Obligation.

\subsection*{Low Rites}
\paragraph{Rite of the Gentle Passing (Low).} Duration: Scene; Range: Touch.  
Materials: a lock of hair from the dying or a drop of water blessed at crossroads.  
Effect: Grant peace to one dying or troubled person. They suffer no additional Harm or stress this scene and may take one final, meaningful action. Gain +1 die to related Comfort/Healing rolls.  
Invoke: 1 action; mark +1 Obligation.  
\emph{Push It:} Their passing leaves a beneficial omen or their words carry supernatural weight; mark 1 SB (Diamonds).  

\paragraph{Rite of the Crossroads Whisper (Low).} Duration: Instant; Range: Self.  
Materials: a stone from a crossroads carried in the mouth for one hour.  
Effect: Gain +1 die to Navigate/Investigate at crossroads (literal or metaphorical). Ask one yes/no question about the ``correct'' path; the Keeper answers truthfully.  
Invoke: 1 action; mark +1 Obligation.  
\emph{Push It:} The path becomes supernaturally clear for one scene, but generate 1 SB (Clubs) as liminal forces take notice.

\subsection*{Standard Rites}
\paragraph{Rite of the Shadow Flock [PASSAGE][VEIL].} Duration: Scene; Range: Near.  
Materials: wool from a black sheep and mist collected at dawn.  
Effect: Summon 3--5 shadow-sheep (Cap 1--2) that can guide through labyrinthine spaces, carry small items, or provide cover [VEIL].  
Invoke: 1 action; mark +1 Obligation.  
\emph{Push It:} One sheep carries a person's shadow for a scene, obscuring them, but mark 1 SB (Spades) as disorientation sets in.

\paragraph{Rite of the Memory Corral [REVEAL][BIND].} Duration: Scene; Range: Zone.  
Materials: a circle of salt mixed with ash from burned letters.  
Effect: Contain lost or suppressed memories. Affected must test Resolve (DV 3) or recall a forgotten memory (Keeper decides). Gain +1 die to related Investigation rolls.  
Invoke: 1 action; mark +1 Obligation.  
\emph{Push It:} Memories gain semi-autonomy and answer directly, but mark +1 Obligation and 1 SB (Hearts).

\subsection*{High Rites}
\paragraph{Rite of the Threshold Covenant [OATH][WARD].} Duration: Extended; Range: Touch.  
Materials: a burned door-frame, a lock of hair from each party, and water blessed by loss.  
Effect: Bind two parties in covenant tied to a threshold. While honored: +1 Effect on threshold actions. Breaking the oath inflicts Harm 2 (Stress/Loss) and blocks crossing for one scene.  
Invoke: Extended ritual; mark +2 Obligation.  
\emph{Push It:} Becomes permanent and geographic (affecting descendants or that place); mark +2 Obligation.  

\paragraph{Rite of the Shepherd's Mercy [CLEANSE][HEAL].} Duration: Scene; Range: Touch.  
Materials: the last breath of someone who died peacefully, kept in unworked silver.  
Effect: Remove one Condition or reduce Harm by two levels. May allow a dying person one final conscious action.  
Invoke: 1 action; mark +3 Obligation.  
\emph{Push It:} Healing is permanent, but the cost transfers---a loved one suffers Harm or calamity strikes; mark 2 SB (Hearts/Clubs).

\subsection*{Obligation}
Starts at 5 for Tier II characters, scaling with tier.

\subsection*{Rivalries}
\begin{itemize}
  \item \textbf{Ikasha (Shadow, Latent Potential):} Subtle friction---Ikasha tends to hidden paths, the Shepherd to endings.
  \item \textbf{The Traveler (Ways \& Roads):} Obvious antagonism---the Traveler opens paths, the Shepherd closes them.
  \item \textbf{Varnek Karn (Necromantic Archives):} Conceptual opposition---Varnek sustains the dead's business, the Shepherd guides to rest.
\end{itemize}

\subsection*{Hard Obligation Triggers (Module-Specific)}
\begin{itemize}
  \item \textbf{Obligation 9+:} The Shepherd appears to demand a service: shepherd a lost soul or memory. Refusal marks you neglectful: worsen Position by one step in related social rolls until penance is made.
  \item \textbf{Obligation 11+:} Connection to liminal spaces destabilizes. Doors may open to the Ways Between, and you risk becoming lost. Requires a quest to restore balance.
\end{itemize}

\subsection*{Melancholy (Persistent Condition)}
Bonding deeply with the Shepherd may grant the \textbf{Melancholy} condition.  
Effect: -1 die to social rolls involving joy or levity, but +1 die to rolls involving comfort, guidance, vigils, or liminal spaces.  
Narrative: A constant reminder of mortality and transition; characters become somber but insightful about endings.

\clearpage
\section{Patron Entry: Morag the Hag}
\label{patron:morag}

\subsection*{Lore}
Morag dwells at the edges of things---crossroads, shorelines, the borders between cultivated land and wild wood. She is ancient beyond measure, her form shifting between that of a wizened crone, a terrible maiden, and a creature of shadows and bone. Her voice carries the weight of countless bargains made and broken, and her laughter echoes with children's tears.  

Morag is the patron of those who seek power through cunning, transformation, and the manipulation of weakness. She offers gifts with hidden curses, teaches transformations that exact a price, and grants knowledge that burns the unworthy. Her realm is perpetual twilight, where hospitality is sacred law and breaking a promise carries penalties worse than death. Her sigil is a twisted knot that shifts between a heart and a noose.  

Those who bargain with Morag gain the ability to see hidden truths, twist fate, and transform themselves or others---but each gift deepens their entanglement in her web of obligations and curses.

\subsection*{Patron's Gift (Imbuement)}
Once per scene as an action (cost: 1 Boon; requires Thiasos), touch an item to imbue it until scene end with +1 Melee and +1 Deception or Insight (when leveraging hidden weaknesses/true nature). \emph{Push It:} extend for one extra scene by marking +1 Obligation.

\subsection*{Low Rites}
\paragraph{Rite of the Twisted Knot (Low).} Duration: Scene; Range: Touch.  
Materials: a rope tied in a complex knot while speaking a name.  
Effect: Bind a minor promise or agreement. Target suffers -1 die on actions that would break the promise. Breaking it generates 1 SB (Hearts) and draws Morag's attention.  
Invoke: 1 action; mark +1 Obligation.  
\emph{Push It:} Breaking causes Harm 1 (Curse) and permanent ``Broken Promise'' Condition; mark 1 SB (Diamonds).  

\paragraph{Rite of the Borrowed Face (Low).} Duration: Scene; Range: Self.  
Materials: a hawthorn mask or convincing portrait.  
Effect: Assume the appearance of someone you know. Gain +2 dice to Deception to impersonate, but cannot mimic memories or mannerisms perfectly. Close acquaintances may test Insight (DV 3).  
Invoke: 1 action; mark +1 Obligation.  
\emph{Push It:} Disguise is perfect, but the person suffers 1 Fatigue, and you may too (Resolve DV 2). Generate 1 SB (Spades).

\subsection*{Standard Rites}
\paragraph{Rite of the Thorned Bargain [OATH][CURSE].} Duration: Extended; Range: Near.  
Materials: blood from each party mixed with thorn-juice over fire.  
Effect: Create binding bargain with enforceable terms. Both parties gain +1 die when fulfilling it. Breaking it inflicts Harm 2 (Curse) and grants +2 dice to the other party's actions for one session.  
Invoke: Extended ritual; mark +1 Obligation.  
\emph{Push It:} Bargain becomes self-enforcing, cruel, and inescapable. Mark +1 Obligation and 1 SB (Clubs).  

\paragraph{Rite of the Bone-Truth [REVEAL][BIND].} Duration: Scene; Range: Near.  
Materials: a bone from a significant event ground into truth-sap.  
Effect: Force one target to answer three questions truthfully. They may resist with Resolve (DV 4).  
Invoke: 1 action; mark +1 Obligation.  
\emph{Push It:} Target can never lie about revealed truths again. Lying inflicts Harm 1. Mark +1 Obligation.

\subsection*{High Rites}
\paragraph{Rite of the Soul's Price [FOLLOW-UP][CURSE].} Duration: Instant; Range: Touch.  
Materials: a possession of the target, their blood, and a loved one’s hair.  
Effect: Inflict curse tied to desire/fear: -2 dice, auto-generate SB, or steal Skill/Talent for one scene.  
Invoke: 1 action; mark +2 Obligation.  
\emph{Push It:} Effect permanent or spreads to loved ones. Mark +2 Obligation and 2 SB.  

\paragraph{Rite of the Cauldron's Boil [TRANSFORM][UNWARD].} Duration: Extended; Range: Zone.  
Materials: a cauldron, symbolic ingredients, and a ``price.''  
Effect: Major transformation: alter a person, location, or magical effect. Permanent, with a price.  
Invoke: Extended ritual; mark +3 Obligation.  
\emph{Push It:} Transformation is perfect but consequences shift elsewhere, or new problems arise. Mark +2 Obligation and start a clock.

\subsection*{Obligation}
Starts at 6 for Tier II characters, scaling with tier. Morag’s obligations are sticky and personal.

\subsection*{Rivalries}
\begin{itemize}
  \item \textbf{Livaea (Temptation, Desire):} Tension---both manipulate desire, but Livaea tempts, Morag bargains.  
  \item \textbf{Aliyah (Curses \& Corruption):} Antagonism---Aliyah corrupts, Morag exacts payment.  
  \item \textbf{Maelstraeus (Infernal Bargainer):} Opposition---Maelstraeus builds webs, Morag leaves one diminished.  
\end{itemize}

\subsection*{Hard Obligation Triggers}
\begin{itemize}
  \item \textbf{Obligation 10+:} Morag appears to collect: demand betrayal, sacrifice, or corruption of someone close. Refusal means she turns against you through manipulation and curses.  
  \item \textbf{Obligation 12+:} Promises misfire, transformations backfire, and others distrust you. Requires a quest to prove worth or risk pariah status.  
\end{itemize}

\subsection*{Bargain-Sight (Persistent Condition)}
Bonding deeply with Morag grants \textbf{Bargain-Sight}.  
Effect: +2 Insight to detect deception or hidden costs. Always see the price of deals, creating paranoia.  
Narrative: Excellent negotiator, but cynical and mistrustful; struggles with genuine bonds.
\clearpage

\subsection{The Inquisitor Prime}
\label{patron:inquisitor-prime}

\paragraph{Lore.}
The Inquisitor Prime is the embodiment of righteous persecution and methodical destruction of the arcane. Neither god nor demon, but a conceptual force born from centuries of witch-hunting. They appear as a figure in pristine white armor stained with the ash of burned spellbooks, their face hidden behind a helm that shifts symbols depending on the viewer's faith.

The Inquisitor values precision over brutality, knowledge over zealotry. Followers are methodical hunters who understand their prey. They gain protection from magical influence, clarity to pierce illusions, and weapons that cut through supernatural defenses. Prolonged service, however, strips away nuance, leaving only intolerance for ambiguity.

Their sigil is a white flame that burns without fuel --- the pure destruction of corruption.

\paragraph{Patron's Gift (Imbuement).}
Once per scene as an action (cost: 1 Boon; requires \emph{Thiasos}), touch an item to imbue it until scene end with \textbf{+1 Melee} and \textbf{+1 Arcana or Insight} (when combating supernatural threats).
\emph{Push It:} Extend one extra scene by marking +1 Obligation.

% --- Low Rites ---
\subsubsection*{Low Rites}

\textbf{Rite of the Pure Flame (Low).}
\emph{Duration:} Scene; \emph{Range:} Self; \emph{Materials:} A candle blessed by a religious authority.
Gain +2 dice to resist supernatural influence (charm, fear, possession) and +1 Armor vs. magical effects. Undead/demonic creatures suffer -1 die to attack you.
\emph{Invoke:} 1 action; mark +1 Obligation.
\emph{Push It:} Extend to allies in Near range, but draw supernatural attention; mark 1 SB (Diamonds).

\textbf{Rite of the Unclouded Eye (Low).}
\emph{Duration:} Scene; \emph{Range:} Self; \emph{Materials:} Holy water mixed with silver.
Gain +2 dice to Insight/Investigate rolls to detect illusions or supernatural deception; cannot be surprised by magical ambushes.
\emph{Invoke:} 1 action; mark +1 Obligation.
\emph{Push It:} Pierce all illusions for one exchange; mark 1 SB (Hearts).

% --- Standard Rites ---
\subsubsection*{Standard Rites}

\textbf{Rite of the Cleansing Light [DISPEL][REVEAL].}
\emph{Duration:} Instant; \emph{Range:} Near; \emph{Materials:} Blessed oil and a silver mirror.
Target a supernatural effect (illusion, enchantment, curse, ward). Test DV = Tier. On Hit: Dispel. On Partial: Weaken. On Miss: Strengthens, generating 1 SB.
\emph{Invoke:} 1 action; mark +1 Obligation.
\emph{Push It:} Completely destroy and trace to its source; mark +1 Obligation and 1 SB (Spades).

\textbf{Rite of the Marked Prey [BIND][CURSE].}
\emph{Duration:} Scene; \emph{Range:} Near; \emph{Materials:} Target’s belonging mixed with blessed salt.
Mark one supernatural target: they suffer -1 die to all rolls and cannot hide with illusion. You gain +1 die to track or strike them. Undead/demonic take +1 Harm from you.
\emph{Invoke:} 1 action; mark +1 Obligation.
\emph{Push It:} Target cannot hide at any distance and suffers Fatigue 1; mark +1 Obligation.

% --- High Rites ---
\subsubsection*{High Rites}

\textbf{Rite of the Consecrated Ground [WARD][BANISH].}
\emph{Duration:} Extended; \emph{Range:} Zone; \emph{Materials:} Relics from three faiths, salt, crushed gems.
Consecrate an area against supernatural influence. Supernatural creatures must test Spirit+Resolve (DV 4) to enter; undead/demonic suffer -2 dice inside. Area becomes [WARD].
\emph{Invoke:} Extended ritual; mark +2 Obligation.
\emph{Push It:} Permanently consecrate a larger area; mark +2 Obligation and start ``Magical Dead Zone [6]'' clock.

\textbf{Rite of the Final Admonition [CLEANSE][FOLLOW-UP].}
\emph{Duration:} Scene; \emph{Range:} Touch; \emph{Materials:} Destroyed spellbook ash, martyr’s water, target’s true name.
Attempt a supernatural execution. Target tests Spirit+Resolve (DV 5). On Fail: destroyed outright. On Success: Harm 3 and -2 dice this scene. Once per target only.
\emph{Invoke:} 1 action; mark +3 Obligation.
\emph{Push It:} Absolute destruction leaves a permanent Sanctified Ground; mark 2 SB (Diamonds).

% --- Patron Traits ---
\paragraph{Obligation.}
Starts at 5 (Tier II), scaling with higher tiers.

\paragraph{Rivalry.}
\emph{Morag (Hag):} Antagonism --- bargains corrupt, Inquisitor purifies.
\emph{Ninth Rim:} Subtle opposition --- erasure vs. revelation.
\emph{Varnek Karn:} Conflict --- preservation of the dead vs. annihilation.

\paragraph{Hard Obligation Triggers.}
\emph{Obligation 9+:} Destroy a source of valued power (weapon, ally, relic) as proof of dedication. Refusal = loss of Inquisitor’s protection.
\emph{Obligation 11+:} Your purity suppresses allied magic; magical healing fails nearby. Must complete a quest to restore balance or become an anti-magic hazard.

\paragraph{Persistent Condition: Inquisitor’s Clarity.}
Those bonded to the Inquisitor may gain the \emph{Clarity} condition.
\emph{Effect:} +2 dice to detect supernatural deception; -1 die to social rolls involving nuance or ambiguity.
\emph{Narrative Weight:} The hunter sees only truth and corruption --- effective at rooting out supernatural foes, but blind to human grayness.

\subsection{Noctira, the Veil of Midnight (Dreams \& Shadows)}

\paragraph{Lore.}
Noctira is the whisper beneath the bed and the flutter of curtains in a still room. She is the half-remembered lullaby sung by the dead, the cold weight on your chest when you cannot move. Her dominion is not over terror itself, but over the boundary between waking and nightmare. Those who follow her learn to weave shadow and dream into both weapon and refuge. Her gifts are seductive: protection in the dark, mastery of dreams, but each step taken draws the supplicant deeper into sleepless nights and hollow-eyed visions.

\paragraph{Patron’s Gift (Imbuement).}
\textbf{Dreamveil:} Once per scene (1 Boon, requires Thiasos), you may blur yourself in shadows. Gain +1 die to Stealth or Resolve rolls until the scene ends. Push It: extend the veil to one ally, but both of you mark +1 Fatigue.

\paragraph{Low Rite.}
\textbf{Rite of the Whispering Shade} — \emph{Scene; Near; Materials: A candle snuffed with your breath.}  
Target suffers -1 die to Courage or Resolve rolls for the scene. Push It: inflict a fleeting hallucination (SB~1) of their worst fear.

\paragraph{Standard Rite.}
\textbf{Rite of Dream-binding [BIND]} — \emph{Extended; Touch; Materials: A name whispered into a dreamcatcher.}  
Target is [BIND]ed by nightmares, unable to recover Fatigue until freed. Push It: Target also suffers a Fear effect (roll on Fear Table, DV~3).

\paragraph{High Rite.}
\textbf{Rite of the Black Lantern [WARD][FOLLOW-UP]} — \emph{Extended; Zone; Materials: A lantern fueled by your blood.}  
Consecrate a location in shadow. Within, you and allies gain +2 dice vs. supernatural fear. Intruders generate +2 SB when attempting to resist dread. Push It: The lantern’s light shows false paths—anyone leaving must resist or become lost in looping halls (advance Location Corruption Clock by 2).

\paragraph{Obligation.}
At 9+: Sleepless visions. You see monsters that may not exist, suffering -1 die to social rolls but +1 die against supernatural foes.  
At 11+: Permanent Condition — \textbf{Hollow-eyed Dreamer}: You exist half in nightmare, suffering -2 dice on mundane tasks but immune to mundane fear.

\section{Haunted House --- ``Midnight Rooms and Shadowed Halls''}
\label{chap:haunted-house}

\begin{tcolorbox}[colback=black!3,colframe=black!40!white,title={Theme \& Atmosphere}]
A haunted house is not merely a place, but a memory made architecture. Its walls bleed with old bargains, its halls echo with unfinished prayers, and its staircases creak under the weight of forgotten sins. To step inside is to invite the midnight in, and once you do, it will know your name.
\end{tcolorbox}

\paragraph*{(Room/Presence/Manifestation)} A hall lined with portraits whose eyes weep brine; a nursery where toys whisper lullabies; a cellar stair that descends farther every night.

\section*{Spades --- Structure \& Rooms}
\label{sec:haunted-house-rooms}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Creaking Hallway} --- Doors never lead to the same place twice.
\item \textbf{Mirror Gallery} --- Reflections move out of sync, sometimes speak.
\item \textbf{Nursery} --- Toys whisper secrets, lullabies warp into screams.
\item \textbf{Ballroom} --- Chandeliers sway to unheard music, floor remembers every dance.
\item \textbf{Library} --- Books rearrange themselves; reading out loud advances clocks.
\item \textbf{Dining Room} --- Feast always set, food rots when touched.
\item \textbf{Cellar Stairs} --- Descend farther each night; bottom never the same.
\item \textbf{Chapel} --- Icons weep blood or saltwater; air tastes of incense.
\item \textbf{Attic} --- Trunks whisper, dolls shift when unlit.
\item[J] \textbf{Garden Court} --- Overgrown vines strangle statuary; paths lead nowhere.
\item[Q] \textbf{Master’s Bedroom} --- Bed still warm; mirror shows stranger watching.
\item[K] \textbf{Basement Door} --- Locked, nailed, chained; something knocks back.
\item[A] \textbf{House-Heart} --- Beating wall, pulsing with shadow; house aware of PCs.
\end{enumerate}

\paragraph*{(Ghost/Resident/Monster)} A grieving mother still searching; a host who insists the dinner continue; a mirror-double who wants your place.

\section*{Hearts --- Residents \& Haunts}
\label{sec:haunted-house-residents}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Widow in Black} --- Weeps into veil; demands someone replace her lost.
\item \textbf{Child-Ghost} --- Plays games with PCs, but stakes are deadly.
\item \textbf{Mirror Double} --- Perfect copy, but hungers for your life.
\item \textbf{Caretaker} --- Keeps house in order, even if it means pruning guests.
\item \textbf{Butler Apparition} --- Polite, precise, cannot be refused.
\item \textbf{Masked Dancers} --- Waltz eternally; join or be trampled.
\item \textbf{Ancestor Shade} --- Claims PCs owe blood-debt for old sins.
\item \textbf{Beast in Walls} --- Snarls, claws, seen only in glimpses.
\item \textbf{Sleepless Scholar} --- Writes endlessly; ink is blood, words become true.
\item[J] \textbf{Lady in Red} --- Arrives at midnight feast, drains warmth with gaze.
\item[Q] \textbf{Cult of the House} --- Mortals who feed it victims in exchange for omens.
\item[K] \textbf{The Returned} --- Dead family member of PC, insists you stay forever.
\item[A] \textbf{Noctira’s Shadow} --- Patron herself manifests; bargain or be claimed.
\end{enumerate}

\paragraph*{(Fear/Threat/Complication)} Lights fail; doors vanish; time loops.

\section*{Clubs --- Complications \& Threats}
\label{sec:haunted-house-complications}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Cold Spot} --- Fatigue 1 on entry; lingering chill seeps in bones.
\item \textbf{Blackout} --- Lights extinguish; SB +1 until rekindled.
\item \textbf{Locked In} --- Doors vanish or lock; only one path remains.
\item \textbf{Time Loop} --- Scene repeats with slight changes.
\item \textbf{Portraits Watching} --- PCs marked with unseen eyes; +1 SB when lying.
\item \textbf{Unseen Hands} --- Objects thrown; tests at DV +1 to act.
\item \textbf{Voice in Ear} --- One PC hears whispers only they can hear.
\item \textbf{Floor Gives Way} --- Collapse; risk Fatigue or Harm.
\item \textbf{Bleeding Walls} --- Scene escalates; Location Corruption +1.
\item[J] \textbf{Phantom Feast} --- Table laid; eat or mark Fatigue.
\item[Q] \textbf{Inverted Room} --- Gravity turns; chandeliers become floor.
\item[K] \textbf{House Divides} --- Party split across alternate layouts.
\item[A] \textbf{Manifest Horror} --- House births a unique foe (see Resident).
\end{enumerate}

\paragraph*{(Secret/Resolution/Desire)} Burial, burning, bargains, bargains.

\section*{Diamonds --- Revelations \& Resolutions}
\label{sec:haunted-house-revelations}
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Hidden Journal} --- Reveals house’s first bargain with Noctira.
\item \textbf{Unmarked Grave} --- Remains must be buried or burned.
\item \textbf{Cursed Painting} --- Destroy to weaken house.
\item \textbf{Secret Passage} --- Leads outside, but costs a memory to open.
\item \textbf{Found Contract} --- House fed by bargains written in blood.
\item \textbf{Silver Key} --- Opens one impossible door; house resents its use.
\item \textbf{Song of Midnight} --- Sung aloud, advances Dread but halts Complication.
\item \textbf{Mirror Bargain} --- Trade reflection for safe passage.
\item \textbf{Family Heirloom} --- Anchors haunting; destroy or restore to resolve.
\item[J] \textbf{House’s Will} --- The house itself writes demands in walls.
\item[Q] \textbf{Noctira’s Bargain} --- Patron offers power for permanent Obligation.
\item[K] \textbf{Blood Price} --- One PC must willingly suffer permanent scar.
\item[A] \textbf{Heart Consumed} --- House awakens fully; either claimed as Patron’s temple or destroyed utterly.
\end{enumerate}

\section*{Quick use notes}
\label{sec:haunted-house-quick}
\begin{itemize}
\item Draw until you have all four suits: Spade = room, Heart = resident, Club = complication, Diamond = revelation. Highest rank sets the main Clock (2--5 $\rightarrow$ 4, 6--10 $\rightarrow$ 6, J/Q/K $\rightarrow$ 8, A $\rightarrow$ 10).
\item Diamonds are codified resolutions (graves, contracts, bargains) that shift position instead of calling for a roll.
\item If any A appears, echo \textbf{house-heart} motifs --- pulsing walls, sentient architecture, bargains offered directly by Noctira.
\end{itemize}

\subsection*{Additional Features}
\begin{itemize}
\item \textbf{Fear Tests:} When confronted by supernatural phenomena, PCs roll Resolve vs. DV 2–4. Failure: roll on Fear Effects Table (see \ref{sec:fear-effects}).
\item \textbf{Location Corruption Clock:} Each Club result adds 1 tick. At full, house collapses into nightmare realm.
\item \textbf{Dread Clock:} Each failed Fear test advances Dread by 1. At max, PC suffers permanent Scar or Corruption.
\end{itemize}

\paragraph{Clock Template.}
\begin{itemize}
  \item \textbf{Dread Clock (10):} PCs’ psychological unraveling.
  \item \textbf{Location Corruption (10):} House becomes increasingly hostile.
  \item \textbf{Resident Sanity (8):} Other occupants succumb to the haunting.
  \item \textbf{Supernatural Escalation (12):} Full manifestation of the haunting.
\end{itemize}

\paragraph{Trigger Examples.}
\begin{itemize}
  \item Mild (+1): Flickering lights, cold spots, phantom footsteps.
  \item Moderate (+2): Apparitions demand attention, objects thrown by unseen hands.
  \item Severe (+3): Entire room shifts, time loop begins, entity manifests.
\end{itemize}

\subsection{Halloween One-Shot: The House Where Midnight Sings}

\paragraph{Premise.}
On All Hallows, a manor on the edge of town wakes. Its windows glow though no one lives there. Noctira’s followers whisper that a bargain lies waiting: the gift of eternal art, if one dares to survive a single night inside.

\paragraph{Setup.}
\begin{itemize}
  \item PCs are lured by promise of wealth, secrets, or patron bargains.
  \item Establish Dread Clock (start at 2/10).
  \item Haunted House Generator: draw 1 card per suit to define core elements.
\end{itemize}

\paragraph{Act I — Invitation.}
\begin{itemize}
  \item The house greets each PC with a room tailored to their fear (advance Dread by +1).
  \item Resident (Heart card) makes first subtle appearance.
\end{itemize}

\paragraph{Act II — Entanglement.}
\begin{itemize}
  \item Location Corruption Clock begins advancing: doors vanish, mirrors show alternate selves.
  \item PCs must investigate to learn what the house wants (Diamond card).
  \item Optional: a Noctira cultist offers aid, at the cost of a boon and a shadow-mark.
\end{itemize}

\paragraph{Act III — Confrontation.}
\begin{itemize}
  \item Complication (Club card) erupts fully.
  \item Residents manifest; confrontation may be combat, exorcism, or bargaining.
  \item PCs must choose: appease, banish, or embrace Noctira’s bargain.
\end{itemize}

\paragraph{Resolution.}
\begin{itemize}
  \item Success: House sleeps, PCs gain haunted boon or art infused with shadow.
  \item Failure: PCs leave scarred, corrupted, or bound as new residents.
  \item Bargain: Accept Noctira’s patronage; gain Dreamveil gift but mark permanent Obligation.
\end{itemize}

\end{document}

