\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{array}
\onehalfspacing

\title{The Shattered Covenant: Black Barrow Keep's Fall}
\author{A Fate's Edge Adventure for Tiers IV-V (Paragon to Mythic)}
\date{}

\begin{document}

\maketitle

\section{Core Premise: The Tragedy That Broke the World}
Centuries ago, \textbf{King Lagan the Unyielding}---a mortal ruler of Payden's Port---commissioned the Aelinnel (gnome) craftsmen of Black Barrow Keep to forge not just servants, but a \textbf{Covenant Heart}: a self-winding engine meant to \textit{guarantee} his lineage's eternal rule through perfect, unbreakable oaths forged in brass and blood. He believed: \textit{Mortal vows fade. Machine oaths endure.}

But the Aelinnel, wary of binding souls to cold perfection, wove a hidden clause into the Heart's gears: \textbf{True oaths require breath, not just tick}. When Lagan activated it, the Heart didn't enshrine his power---it \textit{shattered}. The Covenant wasn't with the king---it was with the \textit{Barrow itself}, a slumbering entity older than men. Enraged, Lagan tried to seize the Heart\ldots{} and the Aelinnel servants, their souls bound to the Barrow's will via the Heart's flaw, turned on him.

What happened next:
\begin{itemize}
  \item The servants didn't just kill Lagan's family---they became \textit{hosts}. As they slaughtered the royal line, whispers slithered into their brass veins: \textit{demons} drawn to the Heart's broken covenant, seeking to wear flesh and steel as new bodies.
  \item Lagan's last act was to seal the upper keep---not to trap threats \textit{in}, but to lock the demon-hosts \textit{out} of the catacombs, where the Barrow's true heart slept.
  \item Now, the Mist has seeped through the seals. The demon-hosts stalk the upper floors, their movements jerky, their eyes glowing with hellfire. Below, the Barrow's dead---kings, servants, oath-breakers---are stirring, kissed by Mist-Wraiths that carry the \textbf{Echo of the Mist King}: a forgotten monarch whose compulsion bends the dead to his will.
  \item \textbf{The Bell Line fails} not from neglect, but because the Mist King's Echo \textit{drowns} the wards' living rhythm. Repairing it requires either:
  \begin{itemize}
    \item (A) \textit{Silencing the Covenant Heart} (stopping the demon threat upstairs), \textbf{AND}
    \item (B) \textit{Facing the Mist King} in the catacombs (to still his Echo or break his hold).
  \end{itemize}
\end{itemize}

\begin{quote}
\textit{The road is a ledger. Every broken wheel is a debt, every lit lamp a witness. The only question is: what are you willing to owe? \\ 
And sometimes\ldots{} the heaviest debt is the one you owe to the king who tried to outlive his mortality.}
\hfill ---\textit{Adapted from Dusana of the Raven Road, The Hearth Ledger}
\end{quote}

\section{Why This Fits Fate's Edge}
\begin{tabular}{|p{0.45\linewidth}|p{0.5\linewidth}|}
\hline
\textbf{Fate's Edge Principle} & \textbf{How This Adventure Embodies It} \\
\hline
Narrative First & Every roll ties to the tragedy: Position shifts reflect whether you're \textit{in step} with the Mist King's Echo (Dominant = you resist his pull; Desperate = his rhythm owns you). Harm from wraiths clears Fatigue---but the mist's chill seeps back faster, creating desperate cycles. \\
\hline
Risk Creates Drama & Magic risks backlash \textit{and} demonic temptation (e.g., a Freecaster's spell might work---but the Covenant Heart tries to \textit{tune} your soul to its rhythm, marking Obligation). Combat risks precise Harm from gear-traps (ignoring armor) \textit{or} soul-sapping touch from the undead. \\
\hline
Meaningful Growth & Tier IV-V choices have legacy weight: Spend XP to fortify a refuge (Asset) \textit{or} train a successor (Follower) who might one day face the Barrow's next awakening. Clearing Obligation to the Barrow requires Acts of Service honoring \textit{imperfection} (e.g., leaving a task unfinished). \\
\hline
Consequences Matter & Failing to silence the Heart doesn't just mean more fights---it advances the \textbf{Demon Clock} (more hosts awaken). Ignoring the Mist King advances the \textbf{Echo Clock} (more dead rise, Barrow's corruption spreads). Success isn't \textit{winning}---it's choosing \textit{what kind of scar you leave on the world}. \\
\hline
Modular Design & Run core mystery (Session 1) $\rightarrow$ mechanical gauntlet (Session 2) $\rightarrow$ catacombs approach (Session 3) $\rightarrow$ climax (Session 4). Layer in magic (Ch. 5), Assets/Followers (Ch. 12), and Legacy Engine (Ch. 19) as XP and story demand. \\
\hline
\end{tabular}

\section{Adventure Structure (4 Sessions)}
\textit{The Mist's advance serves as a dual-layer clock: upstairs, the Demon Clock ticks as hosts awaken; downstairs, the Echo Clock swells as the Mist King's influence grows.}

\subsection{Session 1: The Ashes of Lagan's Line (Investigation \& Stakes)}
\textbf{Hook}: Players arrive in Payden's Port as mist thickens like wet wool. Refugees whisper of \textit{brass-handed things} in the keep's ruins and dreams of a \textit{crown of fog} commanding the dead.

\textbf{Key Locations}:
\begin{itemize}
  \item \textbf{The Salt Docks}: Mist clings to cobblestones, revealing \textit{half-melted brass figurines} (Aelinnel craftsmanship) and \textit{scorch marks} like branding irons.
  \begin{itemize}
    \item \textit{Investigation (Wits+Lore, DV 3)}: Success = find a child's shoe fused to a gear fragment (proof of the massacre). Partial = gain 1 Boon, but notice salt crystals \textit{avoiding} the scorch marks (unnatural repulsion).
    \item \textit{Creative Use}: Spirit+Insight (DV 4) to sense the lingering terror---gain +1 die vs. fear effects here (you \textit{feel} the injustice).
  \end{itemize}
  \item \textbf{The Refugee Camp}: A traumatized Aelinnel tinkerer (missing an eye, replaced with a cloudy glass orb) mumbles:
  \begin{quote}
    \textit{We built the Heart to serve \textbf{life}---not to be a cage for \textbf{hunger}. When the king tried to rip it free\ldots{} the Barrow \textbf{answered}. The servants\ldots{} they weren't \textbf{ours} anymore.}
  \end{quote}
  \begin{itemize}
    \item \textit{Social (Presence+Sway, DV 4)}: Success = learn the Heart's flaw (it \textit{rejects} pure mechanical oaths; needs \textit{breath in the brass}). Failure = he attacks in despair (Harm 1; Desperate Position).
  \end{itemize}
  \item \textbf{The Legate's Hall}: Stone hall where Mistlander elders debate sealing the keep \textit{permanently} (fire, collapse) vs. risking entry.
  \begin{itemize}
    \item \textit{Social (Presence+Command, DV 5)}: Success = convince them to \textit{delay} sealing (gain time to investigate); Failure = they begin preparations, advancing the \textbf{Seal-The-Doom Clock} (if full, they collapse the keep's entrance---trapping anyone inside).
  \end{itemize}
\end{itemize}

\textbf{Mechanical Focus}:
\begin{itemize}
  \item \textbf{Position as Moral Weight}: Near the docks, mist feels \textit{heavy with unresolved guilt}---Spirit-based actions gain +1 die (resonance with the tragedy) but risk +1 SB on failure (the Barrow's sorrow lashes back).
  \item \textbf{Story Beats (SB)}: GM spends SB to reveal tragic details:
  \begin{itemize}
    \item 1 SB: A child's laughter echoes in the mist---then cuts off abruptly.
    \item 2 SB: The scorch marks on the docks form a faint, fading sigil: \textit{Aelinnel knot of broken trust}.
    \item 3 SB: A wraith-like shape flows \textit{uphill} toward the keep---carrying a tiny, tarnished crown.
  \end{itemize}
\end{itemize}

\textbf{Cliffhanger}: At dusk, the Mist parts briefly---revealing Black Barrow Keep's silhouette. From a shattered upper window, a single brass hand \textit{twitches} in the fog\ldots{} then stills.

\subsection{Session 2: The Brass Gauntlet (Upper Floors – Demon-Hosts \& the Covenant Heart)}
\textbf{Hook}: Players choose how to enter:
\begin{itemize}
  \item \textbf{The Salt Causeway} (narrow path; requires Athletics+Survival vs. DV 4 to avoid oil-slicked stones \textit{and} pressure plates that trigger grinding gear-traps).
  \item \textbf{The Refugee's Ladder} (rope ascent through shattered shantytowns; risks drawing hosts to civilians; Social: Command+Presence vs. DV 3 to route civilians safely).
\end{itemize}

\textbf{Key Locations}:
\begin{itemize}
  \item \textbf{The Shattered Atrium}: Once a grand hall, now choked with \textit{demon-hosts}---humanoid figures in tattered livery, joints jerking unnaturally, eyes glowing with sickly green fire. They don't \textit{see}---they \textit{sense rhythm} (heartbeats, breath).
  \begin{itemize}
    \item \textit{Threat}: Harm 2 per hit (their claws are superheated brass); ignores non-magic armor.
    \item \textit{Position Shift}: Flanking isn't just tactical---it's \textit{breaking their sync}. Success = they stutter (lose next action); Failure = they anticipate you (gain +1 die vs. you).
    \item \textit{Weakness}: Loud, irregular noises (e.g., smashing pottery) impose -1 die on their attacks (Disperse their sense of rhythm).
  \end{itemize}
  \item \textbf{The Workshop Corridors}: Aelinnel craftsmanship twisted---workbenches hold \textit{half-assembled hosts}, gears exposed and sparking.
  \begin{itemize}
    \item \textit{Clockwork Trap}: Pressure plates trigger \textit{whirring saw-blades} (DV 3 Athletics to dodge; Failure = Harm 1 + gain 1 Exposure [the Heart notices your rhythm]).
    \item \textit{Creative Combo (Wits+Craft)}: Success = jury-rig a gear-jam to delay a trap (gain 1 Boon for next Athletics roll).
  \end{itemize}
  \item \textbf{The Covenant Heart Chamber}: The throne room's center holds the \textbf{Heart}---a pulsing, brass sphere the size of a man's head, veins of molten silver tracing its surface. Around it, demon-hosts stand rigidly still\ldots{} \textit{waiting}.
  \begin{itemize}
    \item \textit{The Heart's Temptation}: It doesn't attack---it \textit{offers}.
    \begin{itemize}
      \item \textit{Mechanic}: When a player takes an action here, the Heart whispers: \textit{Let me steady your hand. Let me still your fear.}
      \begin{itemize}
        \item \textit{Accept (Spend 1 Boon)}: Gain +2 dice to the action\ldots{} but mark \textbf{1 Obligation segment to the Heart} (track separately).
        \item \textit{Refuse}: No immediate cost, but the Heart's whisper lingers---gain 1 Fatigue (the strain of resisting its \textit{help}).
      \end{itemize}
      \item \textit{Clearing Heart-Obligation}: Requires an Act of Service celebrating \textit{imperfect effort} (e.g., \textit{I deliberately miss this gear-tooth on purpose, letting the scrape be uneven}). Success = clear 1 Obligation segment.
      \item \textit{Silencing the Heart (The Key Step)}: \textbf{Not a single roll}---a multi-step process requiring specific skills:
      \begin{enumerate}
        \item \textit{Understand the Flaw} (Lore+Arcana, DV 5): Success = learn the Heart needs \textit{breath in the brass} to stabilize (i.e., introduce \textit{imperfection}).
        \item \textit{Introduce the Flaw} (Craft+Survival, DV 6): Success = jam a foreign, \textit{organic} object into its veins (e.g., a salt-crystal soaked in refugee tears, a living sprig of mistland moss).
        \begin{itemize}
          \item \textit{Critical Twist}: If the object isn't \textit{truly} imperfect (e.g., a perfect gear-tooth), the Heart \textit{rejects it violently}---gears spew molten brass (Harm 2 to all Near).
          \item \textit{Success}: The Heart seizes---not destroyed, but \textit{stalled}. Demon-hosts jerk violently\ldots{} then collapse into inert brass husks.
          \item \textit{Cost}: The stall is temporary. Every time you \textit{clear Fatigue} via [HEAL] or rest, the Heart gains 1 \textit{``Tick''} (it tries to restart). At 5 Ticks, it \textit{jolts back to life} (advance the \textbf{Demon Clock} by 1 segment; hosts begin to twitch).
        \end{itemize}
      \end{enumerate}
    \end{itemize}
  \end{itemize}
\end{itemize}

\textbf{Mechanical Focus}:
\begin{itemize}
  \item \textbf{Obligation as Temptation}: The Heart's offer isn't free power---it's a \textit{debt} that makes you \textit{more} like it (Ch. 5.3.1 Obligation Overflow = Fatigue). Clearing it requires roleplaying humility (Ch. 15.5.2 Off-Screen Clearing).
  \item \textbf{Harm/Fatigue Roller-Coaster}: Host attacks deal Harm 2 (clearing all Fatigue per Ch. 2.10.2)---but the mist's chill seeps in \textit{faster} after, making recovery harder. Describe: \textit{The host's claws sear your arm---white-hot agony floods your veins. You gasp, not from pain, but from the sudden absence of the bone-deep dread that's clung to you since the docks. Your vision sharpens\ldots{} then the mist's whispers crawl back in, colder than before.}
  \item \textbf{Story Beats (SB)}: GM spends SB to reveal the hosts' tragedy:
  \begin{itemize}
    \item 1 SB: A host's jaw works silently---\textit{forming the word \textbf{sorry}} before seizing again.
    \item 2 SB: Green fire flickers in one host's eyes---revealing a \textit{human iris} beneath the demon-glow.
    \item 3 SB: The Heart's pulse stutters---for 3 seconds, it matches \textit{your} resting heart rate.
  \end{itemize}
\end{itemize}

\textbf{Cliffhanger}: As the last host collapses, the Mist in the chamber \textit{parts}---revealing a staircase spiraling \textit{down} into blackness. From below, a sound like \textit{wet stone grinding on stone}\ldots{} and a voice like wind through tombs: \textit{You silenced my jailers\ldots{} but you did not silence me.}

\subsection{Session 3: The Echoing Deep (Catacombs Approach – Awakening Dead \& the Mist King)}
\textbf{Hook}: The staircase descends into catacombs carved from living stone, walls veined with \textit{frozen mist} that pulses faintly blue. The air tastes of ozone and old blood.

\textbf{Key Locations}:
\begin{itemize}
  \item \textbf{The Whispering Vaults}: Niches hold sarcophagi laced with silver---\textit{not} for decoration, but to \textit{contain}. Lids are slightly askew.
  \begin{itemize}
    \item \textit{Threat}: Mist-Wraiths (wisps of screaming fog) drift between tombs. Touch = Harm 1 [Mental] + gain 1 Fatigue.
    \item \textit{Awakening the Dead}: Roll Wits+Survival (DV 4) to avoid disturbing a sarcophagus.
    \begin{itemize}
      \item \textit{Failure}: Lid bursts open---\textit{animated corpse} rises (Harm 1 per hit; ignores non-magic armor).
      \item \textit{Critical Detail}: These aren't mindless zombies. They move with eerie precision, \textit{turning} to face you as one---\textbf{the Mist King's Echo is syncing them}.
      \item \textit{Creative Use}: Spirit+Performance (DV 4) to \textit{sing a counter-lullaby} $\rightarrow$ wraiths falter for 1 round (advantage on next action).
    \end{itemize}
  \end{itemize}
  \item \textbf{The Hall of Oaths}: A long corridor where walls are etched with \textit{oaths in Aelinnel ledger-script}---names of kings, servants, oath-breakers. The carvings are \textit{fresh}, as if rewritten recently.
  \begin{itemize}
    \item \textit{Social (Presence+Command, DV 5)}: Success = read the oaths' true meaning (they're \textit{not} vows---they're \textit{pleas}: \textit{Remember me. Do not let me be forgotten.}).
    \begin{itemize}
      \item \textit{Success}: Gain +2 dice vs. Mental Harm here (you understand the dead's hunger).
      \item \textit{Failure}: The words seem to \textit{shift} as you read---gain 1 Harm [Mental] (the Echo is probing your mind).
    \end{itemize}
    \item \textit{Obligation Mechanic}: The Barrow tracks your \textit{respect for the dead}.
    \begin{itemize}
      \item \textit{Gain 1 Obligation segment}: If you desecrate a tomb (e.g., loot without words) or speak ill of the dead.
      \item \textit{Clear 1 Obligation segment}: Act of Service honoring the forgotten (e.g., \textit{I press my forehead to this cold lid and whisper the name I imagine they bore}).
    \end{itemize}
  \end{itemize}
  \item \textbf{The Mist King's Audience Chamber}: A vast, domed room. At its center: a throne of black stone, empty\ldots{} but the air \textit{shimmers} where a crown should be. Mist pools here like liquid shadow, coiling and uncoiling in slow, hypnotic patterns.
  \begin{itemize}
    \item \textit{The Mist King's Echo}: Not a ghost---but a \textbf{resonance} in the Barrow's stone, a memory of compulsion so strong it bends the dead to its will.
    \begin{itemize}
      \item \textit{Effect}: All undead in the Barrow gain +1 die to attacks \textit{and} ignore non-magic armor (they move with unnatural unity).
      \item \textit{Social Challenge (Ch. 7.12.2)}: To resist or influence the King:
      \begin{enumerate}
        \item \textit{Phase 1 (Wits+Investigation, DV 4)}: Sense the Echo's rhythm in the mist. Success = gain insight (you feel its \textit{``pulse''}); Failure = gain 1 Harm [Mental] (the Echo stutters your thoughts).
        \item \textit{Phase 2 (Presence+Command, DV 5)}: Attempt to \textit{reason} with the Echo or \textit{defy} its pull.
        \begin{itemize}
          \item \textit{Success (Persuade)}: Gain +3 dice vs. undead effects for 1 hour (the Echo hesitates, confused by your voice).
          \item \textit{Success (Defy)}: Gain +2 dice to all rolls vs. Barrow threats (you stand \textit{off-beat} to its rhythm).
          \item \textit{Failure}: The Echo \textit{locks} onto you---must spend 2 Boons to avoid Harm 3 (your thoughts stutter like a broken record).
        \end{itemize}
      \end{enumerate}
    \end{itemize}
  \end{itemize}
\end{itemize}

\textbf{Mechanical Focus}:
\begin{itemize}
  \item \textbf{Position as Rhythmic War}: At Tier V, Position isn't just survival---it's whether you're \textit{in step} with the Mist King's Echo.
  \begin{itemize}
    \item \textit{Desperate}: \textit{Your breath hitches---\textit{inhale\ldots{} exhale\ldots{} inhale}---matching the mist's pull. You are \textit{becoming} its voice.}
    \item \textit{Dominant}: \textit{You stand \textit{off-beat}. The mist coils, confused. For now, you are not its echo.}
  \end{itemize}
  \item \textbf{Story Beats (SB)}: GM spends SB to reveal the King's tragedy:
  \begin{itemize}
    \item 1 SB: A wraith's shape resolves into a child's face---\textit{mouthing a name you almost recognize}.
    \item 2 SB: The mist forms a fleeting crown\ldots{} then dissolves into weeping.
    \item 3 SB: The throne feels \textit{warm} to the touch---for 3 seconds, like stone left in sunlight.
  \end{itemize}
\end{itemize}

\textbf{Cliffhanger}: As you approach the throne, the Mist parts---revealing \textit{a void} where the crown should be. From it, a whisper like ice cracking: \textit{I was forgotten first. Let the world remember\ldots{} or let it end.}

\subsection{Session 4: The King's Choice (Climax – Repairing the Bell Line)}
\textbf{Hook}: The Mist King's Echo isn't a being to destroy---it's a \textit{wound} in the Barrow's soul. Repairing the Bell Line requires addressing \textit{why} he was forgotten.

\textbf{Key Choices}:
\begin{itemize}
  \item \textbf{Option A: Defy the Echo (Combat)}
  \begin{itemize}
    \item \textit{Mechanic}: Engage the Echo as a \textit{Social Challenge} (Ch. 7.12.2) with escalating stakes:
    \begin{enumerate}
      \item \textit{Round 1 (Wits+Investigation, DV 5)}: Find the Echo's \textit{``weak point''} (a memory of when he was first slighted). Success = gain +2 dice vs. him; Failure = he gains 1 Tick (advances Echo Clock).
      \item \textit{Round 2 (Presence+Command, DV 6)}: Either:
      \begin{itemize}
        \item \textit{(A) Shatter His Will}: Deal \textit{``social Harm''} (Ch. 15.8.3.5 Full Ritual). Success = Echo flickers---advance \textbf{Seal Clock} by 2 segments (Bell Line begins to heal). Failure = he counters---gain 1 Harm [Mental] + Echo Clock advances 2 segments.
        \item \textit{(B) Bind His Will}: Attempt to \textit{contain} him (not destroy). Success = Echo stabilizes---gain +1 die vs. Barrow threats \textit{permanently} (but you mark 1 permanent Obligation to the Barrow; clearing it requires Acts of Service \textit{remembering} the forgotten). Failure = he breaks free---Harm 3 + Echo Clock full (Barrow floods with awakened dead).
      \end{itemize}
      \item \textit{Risk}: If Echo Clock fills \textit{before} you succeed, the Barrow's dead rise \textit{en masse}---you must flee or be overwhelmed (Harm 2 per round until escape).
    \end{enumerate}
  \end{itemize}
  \item \textbf{Option B: Remember the Forgotten (Persuasion)}
  \begin{itemize}
    \item \textit{Mechanic}: Perform an \textit{Act of Service} so profound it addresses the King's core wound:
    \begin{quote}
      \textit{To stand between the forgotten and the remembered, and give voice to the voiceless.}
    \end{quote}
    \begin{itemize}
      \item \textit{Requires}:
      \begin{enumerate}
        \item \textit{Find His Name} (Lore+Investigation, DV 6): Search the Hall of Oaths for a \textit{deleted} oath (scratched over, but traces remain). Success = recover his true name (e.g., \textit{King Vorlag the Unmourned}).
        \item \textit{Give Him Voice} (Presence+Performance or Spirit+Lore, DV 7): Speak his name \textit{into the Mist Pool} in the Audience Chamber, telling one true thing about him (not a lie to flatter---the \textit{truth}, even if painful).
        \begin{itemize}
          \item \textit{Success}: The Mist Pool stills. The Echo doesn't vanish---it \textit{changes}. Gain +3 dice vs. Barrow threats \textit{and} the Bell Line flares to life (mist recedes, wraiths dissipate).
          \item \textit{Failure}: The mist rejects your words---gain 1 Harm [Mental] + the Echo grows hungrier (Echo Clock advances 2 segments).
        \end{itemize}
      \end{enumerate}
      \item \textit{Cost}: Even on success, the Barrow remembers its pain. Clearing its Obligation requires regular Acts of Service: \textit{I tell a story where the hero fails---and I honor the try.}
    \end{itemize}
  \end{itemize}
\end{itemize}

\textbf{Resolution Paths}:
\begin{itemize}
  \item \textbf{The Bell Line Heals (Partial Success)}: You stall the Echo \textit{or} give the King voice---but not both. Mist recedes from Payden's Port\ldots{} but lingers in the Barrow's depths.
  \begin{itemize}
    \item \textit{Cost}: The Barrow remains a weakly sealed wound. Every full moon, mist seeps outward (advance a \textbf{Moon Clock}; if full, repeat Session 3 threats).
    \item \textit{Legacy}: If players retreat, a Follower they saved returns years later, seeking to \textit{finish} the work---offering a Legacy Bond (once/session: gain +1 die when acting to preserve forgotten histories).
  \end{itemize}
  \item \textbf{The Covenant is Renewed (Full Success)}: You both stall the Echo \textit{AND} give the King voice. The Bell Line flares \textit{stronger} than before---mist becomes a \textit{tool}, not a threat (e.g., salt-mist reveals hidden paths).
  \begin{itemize}
    \item \textit{Cost}: The Covenant Heart isn't destroyed---it's \textit{reborn}. It now ticks in time with the Barrow's living breath\ldots{} but demands tribute. Clearing its Obligation requires Acts of Service \textit{honoring imperfect oaths} (e.g., \textit{I swear to try---and I will fail, and I will try again}).
    \item \textit{Legacy}: The Barrow becomes a place of pilgrimage. Scholars come to study its \textit{``living wards''}---and players might be called to defend it from those who would exploit its power.
  \end{itemize}
  \item \textbf{The Barrow Hungers (Failure)}: You anger the Echo \textit{and} fail to remember the King. The Mist King's wrath peaks---he doesn't just awaken the dead; he \textit{unmakes} the Barrow's seals.
  \begin{itemize}
    \item \textit{Cost}: The keep begins to sink into the earth. Players must escape (Chase Scene: Athletics+Acrobatics vs. DV 6 per round; failure = Harm 2 + trapped). Payden's Port is abandoned; the mist claims the ruins.
    \item \textit{Legacy}: Refugees whisper of a \textit{``king who walks beneath the stones''}---and someday, someone will try to wake him \textit{on purpose}.
  \end{itemize}
\end{itemize}

\section{Mechanical Deep Dive: Fate's Edge Integration}
\begin{tabular}{|p{0.3\linewidth}|p{0.3\linewidth}|p{0.35\linewidth}|}
\hline
\textbf{Element} & \textbf{Rule Reference} & \textbf{How This Adventure Uses It} \\
\hline
Obligation & Ch. 5.3.1 (Capacity = Spirit+Presence) & \textbf{Heart's Count} (Ticks from clearing Fatigue) = passive Obligation tracker; \textbf{Mist King's Obligation} (from disrespecting dead) cleared via Acts of Service honoring the forgotten (Ch. 15.5.2). Both mirror Patron mechanics but tied to the Barrow's tragedy. \\
\hline
Harm/Fatigue Roller-Coaster & Ch. 2.10.2 (Taking Harm clears Fatigue) & Wraith/host attacks deal Harm that clears Fatigue---but the Barrow's corruption (mist chill, Echo's pull) seeps back \textit{faster}, creating desperate cycles where players \textit{need} to take a wound to think straight\ldots{} only to feel the dread return stronger. \\
\hline
Position as Narrative Weight & Ch. 7.8.6 (Position scales with Tier) & At Tier V, Position reflects \textit{rhythmic alignment}: \\
 & & \textit{Desperate}: In sync with the Barrow's hunger (you're becoming what it wants). \\
 & & \textit{Dominant}: Off-beat (you resist its pull, but it's exhausting). \\
 & & This makes Position a \textit{storytelling tool}, not just a mechanic. \\
\hline
Story Beats (SB) & Ch. 2.9 (GM spends SB for complications) & SB spent to reveal \textit{tragedy, not just threats}: \\
 & & 1 SB: A detail of the massacre (a child's toy, a lover's last note). \\
 & & 2 SB: A glimpse of the King's humanity (his hands shaking as he signed the Heart's pact). \\
 & & 3 SB: The Echo's true pain (he wasn't evil---he was \textit{terrified} of being forgotten). \\
\hline
Creative Skill Use & Ch. 7.3.2 (GM guidance for creative pairs) & Rewards \textit{Aelinnel ingenuity} and \textit{Mistlander wisdom}: \\
 & & Wits+Craft to jury-rig gear-jams (Session 2) \\
 & & Spirit+Performance to sing counter-lullabies (Session 3) \\
 & & Lore+Survival to find salt in mourning (Session 3 Hall of Oaths) \\
\hline
Legacy Engine & Ch. 19 & Success creates \textit{meaningful legacy}: \\
 & & \textbf{Barrow-Obligation} cleared via Acts of Service celebrating imperfection. \\
 & & Followers inherit bonds to \textit{preserve forgotten stories}. \\
 & & The Barrow itself becomes a potential Asset (if healed) or looming threat (if not). \\
\hline
\end{tabular}

\section{Tier-Appropriate Threats \& Rewards}
\subsection{Threats (Scaled for Tier IV-V)}
\begin{tabular}{|p{0.3\linewidth}|p{0.3\linewidth}|p{0.35\linewidth}|}
\hline
\textbf{Threat} & \textbf{Tier IV Effect} & \textbf{Tier V Effect} \\
\hline
Demon-Hosts (brass-jawed) & Harm 2 per hit; ignores light armor & Harm 3 per hit; ignores \textit{all} non-magic armor \\
\hline
Mist-Wraiths & Harm 1 [Mental] + 1 Fatigue on touch & Harm 2 [Mental] + 2 Fatigue; grants +1 Exposure \\
\hline
Covenant Heart's Temptation & On accepted Boon: +1 Obligation segment & On accepted Boon: +2 Obligation segments \\
\hline
Mist King's Echo & On failed Social: +1 Harm [Mental] & On failed Social: +2 Harm [Mental] + Echo Clock advances \\
\hline
\textbf{Narrative Purpose} & \multicolumn{2}{p{0.65\linewidth}|}{Test if players rely on armor or adapt to \textit{precision} threats (Ch. 20.1.1). Make ``scouting'' risky---reinforces mist as active enemy (Ch. 16.4). Make the ``easy'' power \textit{temptingly costly} (Ch. 5.3.4 Hard Choice). Force players to \textit{engage} the narrative, not just roll dice (Ch. 7.11.3).} \\
\hline
\end{tabular}

\subsection{Rewards (Treasure as Leverage - Ch. 17.2)}
\begin{itemize}
  \item \textbf{Minor}: A \textbf{tear-soaked salt crystal} (once/session: clear 1 Fatigue when performing an Act of Service for the forgotten).
  \item \textbf{Standard}: The \textbf{King's Last Letter} (scorched but readable):
  \begin{quote}
    \textit{Spend 1 Boon to activate: For 1 hour, you gain +1 die vs. Barrow threats AND the Mist King's Echo cannot gain Ticks from you. \\ 
    But: After use, roll d6: on 1, gain 1 permanent Corruption segment (the weight of his regret lingers).}
  \end{quote}
  \item \textbf{Major}: The \textbf{Aelinnel Ledger of Broken Oaths} (massive brass book):
  \begin{quote}
    \textit{Open to a random page: Read an oath the Barrow demands you \textbf{break} to prove your sincerity (e.g., ``Swear you will never return here''). \\ 
    If you break it sincerely: Gain +2 dice to all rolls vs. Barrow threats for 1 day + clear 1 Obligation segment to the Barrow. \\ 
    If you break it flippantly: Gain 1 Harm [Mental] + the Barrow marks you as oath-breaker (advance Echo Clock 1 segment).}
  \end{quote}
\end{itemize}

\subsection{Legacy Seeds (Ch. 19)}
\begin{itemize}
  \item \textbf{If players heal the Barrow (Partial/Full Success)}:
  \begin{itemize}
    \item The Covenant Heart now requires \textbf{regular ``imperfection oaths''} to stay stable (e.g., \textit{I swear to try---and I will fail, and I will try again}). Clearing its Obligation: Act of Service celebrating a \textit{beautiful flaw} (e.g., \textit{I baked bread that burnt---and I shared it anyway}).
    \item A refugee child they saved returns years later, apprenticed to an Aelinnel tinkerer, seeking to \textit{tune} the Heart---not to stop it, but to \textit{dance} with its rhythm.
  \end{itemize}
  \item \textbf{If players fail to heal the Barrow}:
  \begin{itemize}
    \item The Barrow's sinking creates a \textbf{new mist-choked valley} where refugees whisper of the ``Walking King.''
    \item Their Follower (if any) inherits a Legacy Bond: Once/session, gain +1 die when acting to \textit{prevent forgetting} (e.g., visiting a grave, telling a story of the fallen).
  \end{itemize}
  \item \textbf{The Mist King's True Legacy}:
  \begin{itemize}
    \item Even if ``defeated,'' his Echo isn't gone---it's \textit{transformed}.
    \item On Full Success: His presence becomes a \textbf{guardian of memory} (once/session: gain +1 die when acting to honor the forgotten).
    \item On Partial Success: He lingers as a \textbf{warning} (once/session: gain +1 die vs. threats that seek to erase history).
  \end{itemize}
\end{itemize}

\section{GM Guidance: Running the Barrow's Tragedy}
\begin{itemize}
  \item \textbf{Describe the \textit{Rhythm} First}: Before every roll, note the Barrow's pulse:
  \begin{itemize}
    \item \textit{Near salt (upstairs)}: The \textit{faint tick-tick} of stalled gears (safe for now---Heart is stalled).
    \item \textit{In the Whispering Vaults}: Mist coiling in slow, hypnotic \textit{inhale\ldots{} exhale\ldots{}} (sets Position for Mist King's influence).
    \item \textit{Entering the Audience Chamber}: A deep, slow \textit{thoom\ldots{} thoom\ldots{}} like a stone giant's breath.
    \item $\rightarrow$ This sets Position \textit{narratively}: Sync = Desperate (you're in its rhythm); Off-beat = Dominant (you're fighting it).
  \end{itemize}
  \item \textbf{Use SB for Emotional Truth}: Don't just spend SB on monsters---spend it on the \textit{humanity beneath the horror}:
  \begin{itemize}
    \item 1 SB: A demon-host's brass finger twitches---\textit{trying to wipe away non-existent tears}.
    \item 2 SB: The Mist King's Echo forms a fleeting sigh\ldots{} like a man realizing he's been alone for centuries.
    \item 3 SB: A sarcophagus lid lifts just enough to reveal a folded, dust-linen shirt inside (a servant's last gift to his child).
  \end{itemize}
  \item \textbf{Embrace the Temptation of Perfection}: When players consider using the Heart's Boon, describe the \textit{allure}:
  \begin{quote}
    \textit{The Heart's veins glow under your palms---so close to steady rhythm. For one breath, you feel it: the Barrow isn't a tomb. It's a \textbf{promise}. All you have to do is hold this note\ldots{} and never let go.}
  \end{quote}
  Then hit them with the cost:
  \begin{quote}
    \textit{But as you imagine that ease, you see the refugee tinkerer's face---not angry, but \textbf{sad}. She's not begging you to stop. She's begging you to \textbf{remember} why a promise kept through fear is no promise at all.}
  \end{quote}
  \item \textbf{Let the Barrow \textit{Remember}}: If players repeatedly use the same tactic (e.g., jamming the Heart with salt), the Barrow \textit{adapts}:
  \begin{itemize}
    \item After 2 successes, increase DV by 1 (it learns to deflect salt).
    \item After 3 successes, it develops a counter (e.g., gear-teeth now \textit{flick} salt away like water off a duck's back).
    \item $\rightarrow$ This forces players to \textit{innovate}, not just repeat (Ch. 7.3.4 WhentoRoll).
  \end{itemize}
  \item \textbf{The True Victory Isn't ``Winning''}: The Barrow doesn't need to be \textit{destroyed}---it needs to be \textit{understood}. Reward players who:
  \begin{itemize}
    \item Spend time reading the Hall of Oaths' inscriptions (not just looting).
    \item Describe how they \textit{honor} the dead before acting (e.g., \textit{I press my forehead to this lid and whisper the name I imagine they bore}).
    \item Choose to \textit{fail} an oath sincerely (per the Ledger of Broken Oaths) to prove their respect.
  \end{itemize}
\end{itemize}

\begin{quote}
\textit{The road is a ledger. Every broken wheel is a debt, every lit lamp a witness. The only question is: what are you willing to owe? \\ 
And sometimes\ldots{} the heaviest debt is the one you owe to the silence between heartbeats---the space where forgotten kings wait to be remembered. \\ 
Owe it well. Or owe it forever.}
\hfill ---\textit{Adapted from Dusana of the Raven Road, The Hearth Ledger}
\end{quote}

\end{document}
