\documentclass[10pt,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{a5paper, margin=1.2cm, top=1.5cm, bottom=1.8cm}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{titletoc}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[pdfusetitle]{hyperref}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\tcbuselibrary{skins,breakable}

% Color palette from other Fate's Edge documents
\definecolor{bronze}{RGB}{205,127,50}
\definecolor{iron}{RGB}{70,70,70}
\definecolor{parchment}{RGB}{245,240,225}
\definecolor{blood}{RGB}{120,30,30}
\definecolor{deepblue}{RGB}{20,45,80}
\definecolor{recap}{RGB}{230,220,180}
\definecolor{velvet}{RGB}{120,20,50}

% Page style
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE]{\slshape\leftmark}
\fancyhead[RO]{\slshape\rightmark}
\fancyfoot[LE,RO]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% Section formatting
\titleformat{\chapter}[hang]{\Huge\bfseries\color{bronze}}{\thechapter}{20pt}{}
\titleformat{\section}{\Large\bfseries\color{iron}}{\thesection}{15pt}{}
\titleformat{\subsection}{\normalsize\bfseries\color{deepblue}}{\thesubsection}{12pt}{}

% Tcolorbox for sidebars
\newtcolorbox{sidebar}[1][]{colback=parchment, colframe=bronze, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{recapbox}[1][]{colback=recap, colframe=bronze!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{mechanicbox}[1][]{colback=gray!10, colframe=iron!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{tipsbox}[1][]{colback=green!5, colframe=green!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{timersidebar}[1][]{colback=red!5, colframe=red!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{heistbox}[1][]{colback=velvet!5, colframe=velvet!80, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}

% Custom commands for mechanics
\newcommand{\tag}[1]{\textsc{[#1]}}
\newcommand{\sb}{\textsc{sb}}
\newcommand{\dv}{\textsc{dv}}
\newcommand{\xp}{\textsc{xp}}

\begin{document}

\title{\Huge The Bronze Gate \\[0.5cm] \large An Urban Intrigue Adventure for Fate's Edge}
\author{Set down by the fires of Castra Ferrum}
\date{Year of the Iron Harvest}

\maketitle
\thispagestyle{empty}

\tableofcontents
\thispagestyle{empty}
\cleardoublepage

\input{copyright}
\chapter{How to Run This Adventure}
\setcounter{page}{1}

\begin{recapbox}[Previously in Fate's Edge]
Your players have survived the bandit attack on Millhaven (\textit{Blood and Silk}) and brokered peace at the Valvano Ford (\textit{Beyond Millhaven}). They have earned the trust of the Vilikari merchant Tema and may have met Khatun Sarnai (Ykrul chieftain) and Captain Rusk (Black Banner mercenary). They carry a Letter of Introduction to a Vilikari lord and have begun to build a reputation. Now they are drawn into the politics of Castra Ferrum, where old grudges and cursed relics threaten to ignite a new war.

\vspace{0.5em}
\noindent \textit{If your group played the prequel adventure \textbf{The Velvet War of Silkstrand}, the exiles carry more than a letter. They carry the memory of Old Kes and the Silk Coin. They know that the Velvet Court's reach is long, and that Livaea's smile hides a knife. That memory may surface here.}

If your party includes the pre-generated characters from the first adventure, note:
\begin{itemize}
\item \textbf{Kestra} may recognize the curse tablets from her studies in Silkstrand.
\item \textbf{Benjamin} may have served alongside legionaries who spoke of the Bronze Gate.
\item \textbf{Sariel} may have heard of Quick Lena through the Velvet Court.
\item \textbf{Elena} may be called upon to treat victims of the seal's wasting curse.
\item \textbf{Thane} may find the Roman-style wards familiar from Aeler engineering.
\item \textbf{Lyra} may use her diplomatic papers to negotiate with Marcellus.
\item \textbf{Finn} may sense the disturbed Ykrul burial ground beneath the gate.
\item \textbf{Zara} may recognize trade routes and tariff implications.
\end{itemize}
\end{recapbox}

The Bronze Gate is a Tier I--II adventure, suitable for characters who have earned 30--60 \xp. It assumes familiarity with the core resolution loop, Position, \dv, Boons, \sb, and timers. The adventure introduces terrestrial patronage, Roman-style wards, and the weight of military honor. It also draws on the heist mechanics from \textit{The Velvet Touch} and the city procedures from \textit{Violets and Stone}.

\begin{sidebar}[Silk Coin Legacy Tokens]
If your group played \textit{The Velvet War of Silkstrand}, each player character may carry one \textbf{Legacy Token} chosen from the list in \textit{Blood and Silk}. These tokens represent the last gifts of the fallen Silk Coin. If you did not play the prequel, your characters fled Silkstrand with only what they could carry --- but the memory of Old Kes still haunts the alleys of that city.

In this adventure, a Legacy Token may be spent for a one-time +2 dice on a roll that directly honors the Silk Coin's code (keeping a promise, protecting a fellow exile, refusing a bribe). The GM may also allow a token to unlock a hidden clue or reveal a secret passage that only a Silk Coin member would know.
\end{sidebar}

\begin{sidebar}[The Prefect's Wisdom]
Set the \dv low. Let the debt do the work. Every seal, every handshake, every whispered promise is a chain -- and chains, once forged, are not easily broken.
\end{sidebar}

\section{Adventure Background}

Castra Ferrum stands at the northern edge of the province, a walled city built atop a Ykrul burial ground. The Bronze Gate is the only pass through the mountains for three weeks' march. Ecktorian officers, tax farmers, and priest-magistrates hold sway through a web of patronage; a favor given is a debt recorded, a debt recorded is a chain forged.

The Ykrul clans camp beyond the walls, allowed inside only on market days. They sacked the city four times in living memory. Every Ecktorian has a grandfather who fell under a Ykrul arrow. The Vilikari traders who bring grain and iron are looked down upon by the elite -- ``cart-lords'' and ``ledger-lickers'' -- and the Black Banner mercenaries camp in the Field of Broken Standards, selling steel to whoever can pay.

Prefect Gaius Marcellus, commander of the garrison, has recovered a lost relic: the Iron Seal of Varus, the standard of the Ninth Hispania Legion, destroyed by the Ykrul two centuries ago. The seal binds oaths with a wasting curse; any who swears on it and lies sickens and dies. Marcellus intends to use it to force Khatun Sarnai, the Ykrul chieftain, to swear a non-raiding treaty. Sarnai refuses to enter the city, and the seal still carries the old curse -- it harms Ecktorians as well.

The party is hired by Tema, a Vilikari grain merchant, to steal the seal. She means to give it to Sarnai as a gesture of peace, opening the steppe trade route. But Marcellus's rival, Censor Cassia, wants the seal destroyed, and a Black Banner captain named Rusk wants to sell it to the highest bidder -- he has already hired a thief called Quick Lena to take it first.

\subsection{The Curse of Varus}

The Ninth Hispania Legion was betrayed by its own commander, who swore loyalty to a Ykrul war chief and then broke his oath. The legionaries cursed his name as they died, and their standard -- a bronze seal pressed into the wet clay of the treaty -- absorbed their dying breath. Anyone who swears an oath on the seal and lies suffers a wasting sickness: first a cough, then fever, then the slow forgetting of names and faces. There is no cure but the breaking of the seal or the fulfillment of the oath.

The Ykrul believe the seal also curses the ground where it rests. The burial ground beneath the Bronze Gate has been restless since Marcellus brought the seal into the city. Warriors dream of their ancestors weeping.

\section{What the Party Should Know}

At the start of the adventure, the players know the following:

\begin{itemize}
\item The Bronze Gate is the only pass through the mountains. Control of the gate is control of trade.
\item Prefect Marcellus has a relic that can force oaths. He means to use it on the Ykrul.
\item Khatun Sarnai will not meet inside the walls. She demands the seal as wergild for the desecrated burial ground beneath the gate.
\item Tema offers a fortune in silver if the party retrieves the seal for her.
\item A thief named Quick Lena works for Captain Rusk. She is already casing the tribunal.
\end{itemize}

\section{Continuity from Millhaven, the Ford, and the Prequel}

If the party met Mareth in \textit{Blood and Silk} or \textit{Beyond Millhaven}, they may receive a letter at the start of this adventure. Mareth has heard of the seal and warns: \textit{``Oaths sworn on bronze are chains that rust from within. The Traveler bids me tell you: count the clauses. There is always a ninth. The Silk Coin learned that too late.''} This letter can grant +1 die on a single Lore roll related to the seal's history.

If the party earned Sarnai's respect at the Valvano Ford (by mediating fairly or showing courage), she will meet them outside the gate without demanding a tribute. If they offended her, she will require a gift of salt or a story before she speaks.

If the party still carries Tema's Letter of Introduction, they may present it to any Vilikari trader in Castra Ferrum for a +1 die on social rolls in the Vilikari quarter. The letter is worn now, but the seal holds.

\begin{sidebar}[Whispers of the Velvet Court]
A rumor circulates in the Vilikari quarter: Madam Serafine has taken an interest in the Bronze Gate. A Livaean courtesan was seen leaving the Prefect's tribunal at midnight, smiling. The party may recognize the velvet touch of the Court. If they do, they know that Livaea's influence bends oaths and breaks wills. Marcellus may already be compromised.
\end{sidebar}

\chapter{Castra Ferrum -- City Sheet}

\section{Mandate and Crisis}

\begin{itemize}
\item \textbf{Mandate 3:} The Prefect's authority is respected but not loved. Public offices operate smoothly; petitions are heard promptly.
\item \textbf{Crisis 2:} Old grudges simmer. The Ykrul camp outside the walls is a constant reminder of past sackings. The garrison is understrength.
\end{itemize}

\section{Districts}

\begin{longtable}{p{3.5cm} p{5cm} p{4cm}}
\toprule
\textbf{District} & \textbf{Tags} & \textbf{District Effect} \\
\midrule
Praetorium & military, sealed gates, tribunals & +1 Position for Petition (order); Expose starts Desperate \\
Vilikari Quarter & tanneries, warehouses, grain scales & DV -1 for Broker (trade); -1 Position for Sanctify \\
Ykrul Camp (outside) & felt wagons, wolf-banners, burial ground & +1 Position for Smuggle (unguarded paths); -1 Position for Petition (Ecktorian law) \\
Field of Broken Standards & mercenary tents, corrals, forges & +1 Position for Intimidate; -1 Position for Host Feast \\
Archive & scrolls, curse tablets, sealed doors & DV -1 for Research/Expose with proof; -1 Position for Bribe \\
Bronze Gate & portcullis, wall-walk, shrine to Mars & +1 Position for Chase (rooftop); -1 Position for Stealth (guards watchful) \\
\bottomrule
\end{longtable}

\section{Factions and Key NPCs}

\begin{longtable}{p{3cm} p{2.5cm} p{4.5cm} p{3.5cm}}
\toprule
\textbf{Faction} & \textbf{Leader} & \textbf{Goal} & \textbf{Leverage / Strings} \\
\midrule
Ecktorian Tribunal & Prefect Marcellus & Enforce order, control the seal & Military escort, jail keys, legal immunity \\
Censor's Office & Censor Cassia & Destroy the seal, expose Marcellus & Archive access, immunity from prosecution \\
Ykrul Gray Ash & Khatun Sarnai & Protect her people, reopen trade & Safe passage, horse remounts, Ykrul guide \\
Vilikari Traders & Tema & Open steppe route for grain & Trade credit, safe house, forged manifests \\
Black Banner & Captain Rusk & Get paid, get out & Mercenary squad (one scene), battlefield gear \\
\bottomrule
\end{longtable}

\subsection{Prefect Gaius Marcellus}
A wiry, scarred veteran with a voice like a parade ground. He speaks in clipped phrases and never repeats himself. He genuinely believes the seal will bring peace -- and also consolidate his power. His debt timer is his primary mechanic. He has a weakness: his younger brother died in a Ykrul raid, and he has never spoken of it. A PC who learns this (Wits+Insight \dv 4) gains +1 die to social rolls against him.

\subsection{Censor Cassia}
A thin, aging woman with ink-stained fingers. Her left eye is a polished silver disc -- a magical lens that reveals lies. She is Marcellus's political rival and knows the seal's curse is still active. She is dying of a slow fever; destroying the seal is her last act before retirement. She will not admit this, but a PC with Medicine \dv 4 can detect her illness.

\subsection{Khatun Sarnai}
Proud, grieving, hard. Her nephew was killed by a Vilikari caravan. She will not trust Ecktorians, but she respects those who keep their word. She remembers the party from the Valvano Ford (if they mediated fairly). Her camp is guarded by her daughter, a fierce rider named Yelü, who will challenge any stranger to a race before allowing them to speak to Sarnai.

\subsection{``Quick'' Lena}
A Sidhi rogue with mismatched eyes and a nervous laugh. She works for Rusk, not out of loyalty but desperation -- she owes a debt to a Sidhi smuggler named Peyton. She is not evil, but she is fast, and she does not like losing. She has a soft spot for urchins; she left a stolen purse on a beggar's blanket once, and the beggar later saved her from the Watch.

\textit{If the party played the prequel, Lena may confess under pressure: ``I knew the Silk Coin. I was an apprentice -- a coward. I ran when the Court came. I still see Old Kes's face.'' This admission grants +1 die to Sway against her.}

\subsection{Captain Rusk}
A former Ykrul hostage, now leading a Black Banner company. He owes no loyalty to any nation. He will sell the seal to whoever pays best -- but he is haunted by a promise he made to a dying legionary. The legionary's name was Varus, and he was the last survivor of the Ninth Hispania. Rusk carries his badge of rank: a bronze clasp shaped like a gate.

\subsection{Tema}
The Vilikari merchant who hired the party in \textit{Beyond Millhaven}. She is thinner now, her hands stained with grain dust. She has invested everything in this trade route; if the seal is not recovered, she will be ruined. She offers not only silver but a share in her future profits -- a \textbf{Trust} (a party pool of 10 \xp that can be spent on shared assets).

\section{City Timers}

\begin{longtable}{p{4cm} c p{5cm}}
\toprule
\textbf{Clock} & \textbf{Segments} & \textbf{Ticks} \\
\midrule
Heist (Lena) & 6 & End of each downtime scene; when full, Lena steals the seal first \\
Watch Patrol & 4 & Failed Stealth, triggered alarms; when full, guards arrive in 2 rounds \\
Cult of Varus (secret) & 8 & Marcellus's secret cult of oath-breakers; when full, they attempt to sacrifice the party \\
\bottomrule
\end{longtable}

\section{Districts in Play}

When a scene takes place in a district, apply the District Effect. If the party acts in a way that aligns with the district's tags, they gain +1 Position or DV -1 (once per scene). If they act against the tags, the GM gains 1 \sb.

\chapter{Session One -- The Bronze Gate}

\section{Opening Scene -- Tema's Warehouse}

The party meets Tema in the Vilikari quarter, a district of tanneries and cheap taverns. The air smells of lime and wet wool. Tema is thinner than they remember, her hands stained with grain dust. A single lamp burns on a crate that serves as a desk.

\begin{sidebar}[What Tema Offers]
\begin{itemize}
\item 200 silver marks (the party's weight in coin)
\item A \textbf{String: Tema's Trust} -- once per arc, she will smuggle a small object past any checkpoint
\item A Letter of Introduction to a Vilikari lord in the south (a future hook)
\item A share in her future profits (if the party invests in her route, they gain a Trust pool)
\end{itemize}
She needs the \textbf{Iron Seal of Varus} stolen from the Prefect's tribunal. She wants to give it to Sarnai as a peace offering.
\end{sidebar}

If the party asks how she learned of the seal, she admits: \textit{``A Ykrul trader told me. He said the ghosts beneath the gate are weeping. The seal must be returned to the burial ground, or the curse will spread.''}

\section{The Bronze Gate -- A Tour of Factions}

Run this as a social tour. The party can visit each location in any order. No rolls are required unless the players initiate conflict. Use the District Effects as appropriate.

\subsection{The Gate Itself}
The portcullis is bronze-banded iron, etched with the names of every legionary who died defending it. A Censor's clerk keeps the \textit{Ledger of Blood} -- a book recording every soul who passes. The Ykrul call it the Weeping Gate; they say the ghosts scratch the stones at night.

\textbf{Mechanical effect:} The gate is a \tag{WARD} (\textit{Limen Sacrum}). Non-Ecktorians must pay 1 silver or present a \textbf{Bronze Pass} (a String) to cross. Ykrul are not allowed inside except on market days. A PC with the \textit{Shadow's Cloak} from \textit{Blood and Silk} may attempt to slip through with the crowds (Stealth \dv 4).

\subsection{Prefect Marcellus -- The Tribunal}
The tribunal is a fortified hall inside the Praetorium. Marcellus invites the party (if they have a patron or bribe a guard) and offers a job: retrieve a stolen tax tablet from a Vilikari smuggler. Reward: a \textbf{Bronze Pass} (one free crossing). Accepting starts his \textbf{Debt Timer [4]}.

If the party refuses, Marcellus watches them leave. \textit{``You will come back. Everyone comes back to the gate.''}

\subsection{Censor Cassia -- The Archive}
Cassia's office is a labyrinth of scrolls and curse tablets awaiting destruction. She offers a different task: steal a copy of Marcellus's correspondence to prove he is misusing the seal. Reward: an \textbf{Archive Key} (+1 die to Lore rolls in the city) and her protection. She demands a \textbf{String: A Favor to the Censor}.

If the party already has a String with the Censor (from the Valvano Ford, if they exposed a fraud), she offers a second String: \textit{Censor's Confidentiality} (one free immunity from investigation).

\subsection{Khatun Sarnai -- The Ykrul Camp}
Outside the walls, a ring of felt wagons guarded by wolf-banners. Sarnai will not enter the city. She demands the seal as wergild for the desecrated burial ground. If the party swears an oath to bring it to her, she offers a \textbf{Ykrul Guide} (a Cap 2 follower who knows the steppe). Her daughter Yelü challenges a PC to a horse race; winning impresses Sarnai and grants +1 die to negotiations.

If the party helped Sarnai at the Valvano Ford (by mediating the blood-price), she already respects them. She will not demand an oath, but she will ask: \textit{``Will you do for my people what you did for the ford?''}

\subsection{Captain Rusk -- The Field of Broken Standards}
Rusk's camp is a ragged collection of tents and corrals. He is friendly but shrewd. He has already hired Quick Lena to steal the seal. He offers the party a cut if they help -- or a bribe to stay out of his way. He has his own \textbf{Debt Timer [4]} (he owes a smuggler named Peyton). Paying his debt wins a temporary alliance.

If the party met Rusk at the Valvano Ford, he remembers them. \textit{``You talk well. Can you fight?''}

\section{End of Session One -- The Heist Timer }

Start a \textbf{Heist Timer [6]}. Each downtime scene (or at the end of each session), tick it +1. When it fills, Quick Lena steals the seal first -- the rest of the adventure becomes a chase.

The party should choose at least one patron by the end of the session. If they cannot decide, Tema offers a small retainer (10 silver each) to keep them interested.

\chapter{Session Two -- The Tribunal Heist}

\section{Opening Scene -- The Night Before}

The party stakes out the tribunal. They can bribe a guard (Presence+Streetwise \dv 3, cost 1 Boon or 10 silver), steal a patrol schedule (Wits+Stealth \dv 4), get a map from Cassia (requires her favor), or wait for a storm (the guards are distracted, but the storm wakes the curse).

Start a \textbf{Watch Patrol Timer [4]}. Each failed Stealth roll ticks it. When full, the alarm is raised.

\begin{timersidebar}[Timer Triggers]
\textbf{Watch Patrol Timer advances +1 when:}
\begin{itemize}
\item A Stealth roll fails.
\item A trap's \tag{ALARM} triggers.
\item A guard is not silenced before they can shout.
\end{itemize}
\textbf{Heist Timer advances +1 when:}
\begin{itemize}
\item A downtime scene ends.
\item The party spends a full hour debating without action (GM's call).
\item A partial success on a critical roll creates a delay (e.g., the lockpick breaks).
\end{itemize}
\end{timersidebar}

\section{Using Heist Mechanics from The Velvet Touch}

Before the heist, each player describes one preparation (casing, bribing, stashing gear). Each strong success on a Setup roll (Wits+Stealth or Wits+Streetwise, \dv 3) grants \(+1\) Position or \(-1\) \dv later. Each player starts with 2 \textbf{Kit Tokens} (lockpicks, dark lantern, false papers) that can be spent during the heist for:

\begin{itemize}
\item +1 die on a roll
\item Improve Position by one step
\item Tick the Heist Timer forward by one segment
\item Pause a Threat Timer (prevent it from advancing for one tick)
\end{itemize}

After use, the token is consumed.

\subsection{Quick Lena's Preparations}

Lena has already cased the tribunal. She knows the patrol schedules and has bribed a kitchen slave. She will enter through the sewers, bypassing the front gate. Her Heist Timer starts at 3 segments (she is halfway done). The party can intercept her if they act quickly.

\section{Chamber One -- The Armory}

\textbf{Ward:} \tag{ALARM} + \tag{ENTANGLE}. Lead sheets on the floor trigger a bell and a net.

\textbf{Counter:} Wits+Tinker \dv 4 to disable. On a partial, the bell rings but the net fails -- tick Watch Patrol +1. A PC with the \textit{Unlocked Window} talent may declare they prepared for this (spend 1 Kit Token) and bypass the trap entirely.

\textbf{Optional:} A curse tablet hidden in the wall (Wits+Investigation \dv 5). If found, it reveals Marcellus has been using the seal to curse rivals. This is a \textbf{String: Marcellus's Treason}. The tablet is written in legionary shorthand; a PC with Lore 2+ can translate it.

\section{Chamber Two -- The Scriptorium}

\textbf{Ward:} \tag{WARD} (\textit{Limen Sacrum}). The door burns non-legionaries.

\textbf{Counter:} Recite a legion password (Wits+Lore \dv 3) or offer a coin to the \textit{Lares} (1 Boon or 1 silver). Failure costs 1 Fatigue; the guard beyond hears the cry. A PC with the \textit{Shadow Meld} talent (Ikasha) may step through the door's shadow without triggering the ward.

Inside: a map of the city marked with Ykrul trade routes, and a forged census tablet showing Marcellus skimming grain tithes -- another String.

\section{Chamber Three -- The Reliquary}

The seal rests in a lead-lined chest on an altar to Mars. Quick Lena is already there, trying to crack the lock.

\textbf{Ward:} \tag{CURSE}. Anyone who opens the chest without a blessing gains \textit{Marked by Mars} -- enemies gain +1 die to track them.

\textbf{Counter:} Pray to Mars (Spirit+Lore \dv 3) or smash the curse tablet hidden under the altar (Wits+Investigation \dv 4, costs one round). A PC with the \textit{Varus's Shard} (from a previous encounter) may use it to absorb the curse instead.

\subsection{Confronting Quick Lena}

Lena is not a fighter. She will negotiate: split the reward, or she will walk away. A Presence+Sway \dv 3 convinces her. On failure, she uses \textit{Always Another Exit} -- a smoke bomb and escape -- but drops her lockpicks (a Minor Asset: +1 die to open locks). Tick Watch Patrol +1.

If the party fights, Lena has Harm 2, Stealth 3, Subterfuge 2. She uses \textit{Light Fingers} to steal a weapon mid-combat (Wits+Subterfuge vs PC's Wits+Notice). At Harm 2 she surrenders. If the party spares her, she offers a \textbf{String: Lena's Gratitude} (once, she will warn them of a rival's scheme).

\subsection{Lena's Secret}

If the party asks why she works for Rusk, she admits: \textit{``I owe Peyton a debt. He paid for my passage out of Galanina. Now he wants the seal. I don't know why.''} This is a hook for a future adventure involving Sidhi smugglers.

\section{Escape -- Skill Challenge}

The party must exit before the Watch Patrol timer fills. Run this as a \textbf{skill challenge}: the players need \textbf{3 successes before 2 failures}. Each PC describes how they contribute (Stealth to avoid patrols, Command to give orders, Deception to bluff past a guard, Athletics to climb a wall). The GM sets \dv based on the approach (typically 3).

Complications on failure:

\begin{itemize}
\item A guard spots the party -- tick Watch Patrol +1. The PC may spend 1 Boon to create a distraction and avoid the tick.
\item A door is locked that wasn't before -- the party must find another route (adds 1 segment to the challenge).
\item Rusk's mercenaries arrive (if the Heist Timer is at 4+ segments). The party must fight or bribe them.
\item A curse tablet activates (if the party disturbed it). The affected PC gains \textit{Varus's Whispers} ( --1 die to resist persuasion) for one day.
\end{itemize}

If the timer fills, run a brief chase. On failure, they are arrested -- Cassia may bail them out in exchange for the seal, or Marcellus may offer a deal.

\chapter{Session Three -- The Curse Unsealed}

\section{Opening Scene -- The Seal's Whisper}

The seal still carries the old curse. Anyone who sleeps near it must make Spirit+Resolve \dv 4 or gain \textit{Varus's Whispers} ( --1 die to resist persuasion). The curse worsens each night. On the second night, the affected PC begins to forget names; on the third, they forget their own face.

Rusk and Lena (if she escaped) are hunting them. Marcellus has declared them outlaws. The party must decide where to hide: the Vilikari quarter (Tema can shelter them, but she may be watched), the Ykrul camp (Sarnai offers protection, but the cold is harsh), or the Archive (Cassia hides them among the curse tablets, but the tablets whisper at night).

\section{The Final Confrontation}

The location depends on the party's chosen patron:

\begin{itemize}
\item \textbf{The Bronze Gate} (siding with Sarnai). Rusk's mercenaries attack at midnight. The party must defend the gate while a Ykrul shaman performs a ritual to purify the seal.
\item \textbf{The Archive} (siding with Cassia). She performs a ritual to destroy the seal. Marcellus arrives with guards. The party must hold them off until the ritual completes.
\item \textbf{The Praetorium} (siding with Marcellus). He forces the party to bind Sarnai's oath. The curse backfires. The party must contain the backlash while Marcellus's cultists try to seize the seal.
\end{itemize}

\section{Breaking the Curse -- Three Ways}

\subsection{Return the Seal to the Ykrul}
Give it to Sarnai. She buries it under a cairn with a steppe ritual. The curse ends. Sarnai opens the trade route. Marcellus becomes a permanent enemy. \textbf{Cost:} The party loses any favor with Marcellus; his Debt Timer becomes a Hunt Timer [6]. The Ykrul offer a \textbf{String: Sarnai's Oath} -- once, they will ride to the party's aid.

\subsection{Destroy the Seal with Fire}
A ritual (Spirit+Arcana \dv 6). On success, the seal shatters. On a partial, it shatters but the curse transfers to the one who strikes it -- that PC gains \textit{Varus's Scourge} (1 Fatigue each dawn until they atone). Cassia rewards a \textbf{Censor's Seal} (one free legal pardon). \textbf{Cost:} Marcellus becomes a permanent enemy; Rusk may demand compensation. The Archive gains a permanent \tag{WARD} against curses.

\subsection{Bind the Seal with a New Oath}
Swear a treaty between Marcellus, Sarnai, and Tema. Requires a \textbf{Persuasion Timer [6]} for each faction. On success, the curse transforms into a blessing -- the gate harms no one. The party gains a String with all three factions. Marcellus still owes them a debt (his Debt Timer remains, but now it's a favor they can call). \textbf{Cost:} The party must mediate a tense ceremony; any failed roll during the oath-swearing triggers a complication (e.g., a Ykrul warrior challenges the prefect to a duel). The party may also gain a \textbf{Trust} of 10 \xp from Tema's gratitude.

\subsection{Secret Fourth Way -- The Ninth Clause}

If the party uncovered the curse tablet and translated the legionary shorthand, they discover that the seal has a ninth clause: \textit{``If the one who swore the oath speaks the name of the betrayed aloud, the curse will devour the betrayer instead.''} The betrayer was the Ninth Legion's commander, long dead. But his ghost walks the burial ground beneath the gate. The party can summon him with a ritual (Spirit+Lore \dv 5). If they convince him to speak his own name, the curse ends forever, and the seal becomes inert. This requires a Presence+Sway \dv 5; failure means the ghost attacks (Cap 3). Success earns a \textbf{Legendary String: The Ninth Legion's Gratitude} -- the ghosts of the legion will warn the party of one ambush, anywhere, at any time.

\section{Ending -- The Gate Opens}

At dawn, the portcullis rises. Describe the scene: the Ykrul riders waiting, the merchants with their carts, the legionaries saluting. Then ask each player: \textit{``What does your character want now? And what are you willing to owe to get it?''}

If the party succeeded, Tema gives them each a silver medallion stamped with a gate. \textit{``Show this to any Vilikari trader. They will know you are trusted.''}

If they failed, they escape with their lives but without the seal. Rusk now holds it. He may sell it to the highest bidder -- or Peyton may call in his debt. The gate remains closed. The Ykrul grow restless. War may come before spring.

\begin{tipsbox}[Continuing the Campaign]
The party now has enemies (Marcellus, perhaps Rusk) and allies (Sarnai, Cassia, Tema). Their reputation precedes them. Future adventures could involve:
\begin{itemize}
\item Marcellus seeking revenge through legal or military means.
\item A Ykrul uprising sparked by the broken curse.
\item Rusk's debt to Peyton coming due -- and the party being dragged into Sidhi smuggling politics.
\item The Censor calling in her favor to expose a larger conspiracy.
\item A Sidhi ghost ship appearing in the harbor, carrying a message from Quick Lena's past.
\item The ghosts of the Ninth Legion asking the party to find their lost eagle standard.
\item The Velvet Court, having lost interest in the seal, now sets its sights on the party's new home.
\end{itemize}
\end{tipsbox}

\chapter{Advancement and Rewards}

\section{Experience Points (8--12 \xp per PC)}

\begin{itemize}
\item Attendance: +2
\item Retrieve the seal: +2
\item Break or secure the curse (any method): +2
\item Neutralize Quick Lena (negotiation or combat): +1
\item Gain a String from two different factions: +1 per faction (max +2)
\item Complete a Terrestrial Patron's Debt Timer : +1
\item Personal goal (from player's own backstory): +1--2
\item Discover the Ninth Clause: +1
\end{itemize}

\section{Strings and Assets}

\begin{itemize}
\item \textbf{Bronze Pass} -- One free crossing of the Bronze Gate. May be gifted to another.
\item \textbf{Archive Key} -- +1 die to Lore rolls in Castra Ferrum. Also opens the sealed curse tablet vault.
\item \textbf{Sarnai's Favor} -- Ykrul safe passage for one journey. Includes a remount at any Gray Ash camp.
\item \textbf{Marcellus's Debt [4]} -- Track this timer. When full, he demands a dangerous service. If the timer is cleared by service, the party gains a \textbf{Prefect's Pardon} (one free arrest immunity).
\item \textbf{Censor's Seal} -- One free legal pardon (destroyed after use). Also grants access to the Censor's private library.
\item \textbf{Varus's Shard} -- A piece of the broken seal. Once per arc, curse a target with --1 die to all rolls for one scene; user also takes 1 Fatigue. If the shard is buried under a Ykrul cairn, it loses the curse and becomes a \textbf{Talisman of the Ninth} (+1 die to resist oath-breaking).
\item \textbf{Tema's Trust} -- Once per arc, she will smuggle a small object past any checkpoint. Also grants a 10\% discount on Vilikari goods.
\item \textbf{Ykrul Guide} (Cap 2 follower) -- Knows the steppe, speaks Ykrul and Ecktorian, can find water in dry country.
\end{itemize}

\chapter{Appendix -- Quick Reference}

\section{Timer Summary}

\begin{longtable}{p{4cm} c p{5cm} p{3cm}}
\toprule
\textbf{Clock} & \textbf{Segments} & \textbf{Ticks} & \textbf{When Full} \\
\midrule
Heist (Lena) & 6 & End of each downtime scene & Lena steals the seal first -- chase begins \\
Watch Patrol & 4 & Failed Stealth, triggered alarms & Guards arrive in 2 rounds \\
Cult of Varus & 8 & PC actions that empower Marcellus & Cult attempts to sacrifice the party \\
\bottomrule
\end{longtable}

\section{Trap Tags (Roman Flavor)}

\begin{longtable}{p{2.5cm} p{3cm} p{4cm} p{3cm}}
\toprule
\textbf{Tag} & \textbf{Roman Name} & \textbf{Effect} & \textbf{Counter} \\
\midrule
\tag{ALARM} & \textit{Vigilum Signum} & Bell rings, ticks Watch Patrol & Stealth, cut thread \\
\tag{ENTANGLE} & \textit{Plumbum Manus} & Net restrains & Athletics \dv 3 to break \\
\tag{WARD} & \textit{Limen Sacrum} & Burns non-legionary & Offer coin, recite password \\
\tag{CURSE} & \textit{Tabula Defixionis} & Inflicts condition & Smash the tablet, fulfill the oath \\
\bottomrule
\end{longtable}

\section{NPC Quick Stats}

\subsection{Quick Lena -- Rogue (Tier II)}
Wits 3, Stealth 3, Subterfuge 2. Talents: \textit{Light Fingers}, \textit{Always Another Exit}, \textit{The Ghost Did It} (once). Harm 2. Secret: she owes a debt to Peyton, a Sidhi smuggler.

\subsection{Prefect Marcellus -- Terrestrial Patron (Tier III)}
No combat. Offers Debt Timer . Leverage: guards, jail, military escort. Secret: he leads a cult of oath-breakers who worship Varus's ghost.

\subsection{Censor Cassia -- Sage (Tier II)}
Wits 4, Lore 3, Arcana 2. Can see lies through her silver eye (automatically detect Deception \dv 4 or less). Secret: she is dying; destroying the seal is her final act.

\subsection{Khatun Sarnai -- Warrior (Tier II)}
Body 3, Melee 3, Command 2. Tracks honor debts separately. Secret: her daughter Yelü is the child of a legionary who defected.

\subsection{Captain Rusk -- Mercenary Captain (Tier II)}
Body 3, Melee 3, Command 2. His personal Debt Timer [4] is owed to the smuggler Peyton. Secret: he carries Varus's badge of rank.

\subsection{Tema -- Vilikari Merchant (Tier I)}
Presence 2, Sway 2, Commerce 2. No combat. Offers Trust and trade discounts.

\section{City Actions (from Violets and Stone)}

When in Castra Ferrum, PCs may use these urban actions:

\begin{itemize}
\item \textbf{Grease Palms (Wits+Streetwise)}: Secure permits or toll waivers. Hit: tag \textit{Greased Palms}; miss: rumor of bribery (Exposure +1).
\item \textbf{Audit Ledger (Wits+Investigation)}: Catch graft or tithe anomalies. Hit: DV -1 on later Expose.
\item \textbf{Carouse for Rumors (Presence+Performance)}: Trade coin and stories. Hit: bank 1 Clue; miss: Hangover Debts (GM gains 1 \sb).
\item \textbf{Forge Documents (Wits+Tinker)}: Create false papers. Hit: tag \textit{Forged Papers}; miss: Exposure +1.
\item \textbf{Call the Watch (Presence+Command)}: Leverage order. Hit: \textit{Watchful Patrols} tag; miss: they arrive biased.
\end{itemize}

\chapter{Colophon}

This account was set down in the Prefect's own scriptorium, by lamplight, on paper seized from a Vilikari caravan. The ink is lamp-black and the ash of a burned treaty. The binding is legionary leather, stained with rust.

The gate still stands. The debts remain unpaid.

--- Set down by a clerk of the Bronze Gate, Year of the Iron Harvest

\chapter*{Appendix: The Gate Remembers --- A Letter from Mareth}
\addcontentsline{toc}{chapter}{Appendix: The Gate Remembers --- A Letter from Mareth}
\markboth{APPENDIX: THE GATE REMEMBERS}{APPENDIX: THE GATE REMEMBERS}

\begin{quote}\itshape
To those who walked the broken road and crossed the ford:

I hear you have come to the Bronze Gate. A threshold of iron and old ghosts. The Traveler whispers that the seal you seek is not just bronze --- it is a promise that was never meant to be kept. I have seen such things before, in the vaults of Thepyrgos. The wise do not break such promises. They fulfill them, or they forget them, or they find a clause that the dead did not foresee.

I was there when the Silk Coin burned. I pulled Kestra from the canal. She does not remember me. The living forget. The dead do not.

I am not there to guide you. But I can offer what I have learned since we parted. The road has taught me that magic is not a tool. It is a conversation. In a city of ledgers and oaths, every word you speak is a thread. Every silence is a knot. Choose carefully.

I have enclosed a few reflections. Use them as you will.

\textit{Walk true, but not alone.}

Mareth
\end{quote}

\section*{Magic in the Shadow of the Gate}

Castra Ferrum is not like the open steppe or the broken marches. It is a place of walls, oaths, and recorded debts. Magic here draws the attention of the Censor's silver eye, and the Prefect's clerks note every whispered prayer to an unknown patron. Yet the city is also a threshold --- built on a burial ground, sealed by a cursed relic, guarded by ghosts who remember every broken vow.

A wise spellcaster adapts. The following advice applies to any character who uses magic (Runekeepers, Cantors, Invokers, or free casters) in this adventure.

\subsection{On Stealth and Sorcery}

Magic is not silent unless you make it so. In the tribunal, a flash of light or a sudden gust of wind will raise the Watch Patrol timer. Consider:

\begin{itemize}
\item \textbf{Subtle tags} (from threadweaving) or \textbf{Ikasha's rites} can muffle sound and bend shadows.
\item \textbf{Offering a coin to the Lares} before casting a Rite near a \tag{WARD} may placate the household spirits and prevent alarm.
\item \textbf{Casting during a storm} (the wind hides whispers) grants +1 Position to Stealth rolls for magic, but the GM gains 1 \sb (the storm wakes the dead).
\end{itemize}

\begin{sidebar}[Mareth's First Rule]
\textit{``Never cast a spell you cannot explain as coincidence. A lock that opens by itself is a theft. A guard who falls asleep mid-sentence is an investigation. But a candle that gutters at the wrong moment? A gust of wind that blows out the lantern? Those are accidents. Let the Watch believe in accidents.''}
\end{sidebar}

\subsection{On Curses and Oaths}

The Iron Seal is not a normal magical object. It is a \textit{broken oath made solid}. Any magic that touches it --- divination, warding, even telepathy --- risks awakening the curse prematurely. The GM may spend \sb to have the curse leak out: a PC who rolls a 1 on a spellcasting roll near the seal gains 1 Fatigue (the ghost of Varus whispers their name) and the curse's effects worsen by one step.

To protect themselves, a caster may:

\begin{itemize}
\item \textbf{Swear a minor oath to Mars} (the legion's patron) before handling the seal. This grants +1 die to resist the curse for one scene, but marks a \textbf{Debt Timer [4]} to the war god (fulfill a violent task or lose the blessing).
\item \textbf{Use a \textit{Tabula Defixionis} (curse tablet) as a focus}. A PC who finds one of Cassia's tablets may redirect the seal's backlash into the tablet. The tablet shatters, but the curse does not affect the caster.
\item \textbf{Invoke the Traveler's \textit{Rite of the Forked Sign}} at the Bronze Gate's threshold. This creates a temporary \tag{WARD} against the curse, but the sign is visible to anyone with arcane sight (including Cassia's silver eye).
\end{itemize}

\subsection{On Negotiation and Enchantment}

The social magic of Livaea (the Velvet Invitation, the Unrefusable Offer) is powerful in a city of patrons and clients. But Castra Ferrum's elite are paranoid. A guard who suddenly agrees to a bribe may later remember the encounter as unnatural.

To avoid suspicion:

\begin{itemize}
\item \textbf{Always pair enchantment with a plausible mundane reason}. \textit{``He let us through because we reminded him of his daughter''} is believable. \textit{``He let us through for no reason''} is not.
\item \textbf{Spend a Boon to layer a false memory}. After a successful Livaean Rite, the target recalls a perfectly ordinary conversation. The GM gains 1 \sb (the lie will surface if investigated).
\item \textbf{Do not use the same Rite twice on the same faction}. Marcellus's clerks compare notes. A second unnatural charm ticks the Cult of Varus timer.
\end{itemize}

\subsection{On the Dead Beneath}

The Ykrul burial ground is not empty. The ghosts of the Ninth Legion and the Ykrul warriors who killed them still walk at midnight. Magic that disturbs them --- divination, summoning, even a simple \textit{Cradle Song} to put a guard to sleep --- may draw their attention.

\begin{itemize}
\item \textbf{If a PC rolls a 1 on any magic roll at night}, the GM may have a ghost appear. It is not hostile. It asks a question: \textit{``Why do you walk on our bones?''} Answering truthfully (and respectfully) grants +1 die on the next magic roll. Lying or attacking advances the Cult of Varus timer by +1.
\item \textbf{A character with the \textit{Beast-Tongue} talent} may hear the horses in the Ykrul camp speak of the ghosts. The horses do not lie, but they speak in images, not words.
\item \textbf{A Cantor who sings the \textit{Bridge's Witness}} at the gate may hear the voices of the dead legionaries. They do not answer questions, but they may sing a verse of the Ninth Legion's oath --- which is also the key to the Ninth Clause.
\end{itemize}

\section*{New Low Rites for the Traveler (From Mareth's Letter)}

These rites are taught by the letter alone. A character who has learned the Traveler's low rites from Mareth may spend a downtime scene studying them; no XP cost is required, but the character must be in a liminal space (a crossroads, a bridge, a gate) to attune.

\begin{sidebar}{Rite of the Unsealed Gate (Low Rite)}
\textbf{Cost:} 1 action, mark +1 Obligation. \\
\textbf{Effect:} Touch a locked door, gate, or chest. For the next minute, the lock's mechanism is muffled --- any attempt to pick it is +1 Position, and the lock makes no sound when it opens. If the lock is warded, this Rite suppresses the \tag{ALARM} tag for one exchange. \\
\textbf{Push:} Mark 1 Fatigue to extend the effect to all locks in the same room (up to three). \\
\textit{Mareth's note: ``The gate wants to open. You just have to ask quietly.''}
\end{sidebar}

\begin{sidebar}{Rite of the Witnessed Step (Low Rite)}
\textbf{Cost:} 1 action, mark +1 Obligation. \\
\textbf{Effect:} You lay a small offering (a coin, a hair, a drop of blood) at a threshold. For the next hour, any creature that crosses that threshold while intending to deceive you must roll Resolve \dv 3 or be compelled to speak a single true word aloud (GM chooses the word based on context). The offering is consumed. \\
\textbf{Push:} Mark 1 Fatigue to make the effect last until dawn. \\
\textit{Mareth's note: ``The Censor uses a version of this. Her silver eye sees more than she admits.''}
\end{sidebar}

\begin{sidebar}{Rite of the Debt Reminder (Low Rite)}
\textbf{Cost:} 1 action, mark +1 Obligation. \\
\textbf{Effect:} Name a person you have a String on (a favor owed, a secret held). For the rest of the scene, they are visibly uneasy in your presence --- they suffer --1 die to Deception against you, and you gain +1 die to Sway when reminding them of the debt. This Rite does not create a new String; it only amplifies an existing one. \\
\textbf{Push:} Mark 1 Fatigue to extend the effect to all Strings you hold on the same faction. \\
\textit{Mareth's note: ``Debt is a chain. This Rite is the rattle of the links.''}
\end{sidebar}

\section*{On the Patrons of the Gate}

Castra Ferrum has its own patrons, old and hungry. The party may encounter signs of them:

\begin{itemize}
\item \textbf{Mars, the War God:} His shrine at the Bronze Gate is tended by a one-legged priest who offers blessings to those who swear to fight for the city. Accepting his blessing grants +1 die to melee combat for one scene, but marks a \textbf{Debt Timer [4]} to Mars (a future battle must be fought in his name).
\item \textbf{The Lares (Household Gods):} Every Ecktorian home has a small shrine. Offering a coin or a pinch of salt before a heist grants +1 Position on the first Stealth roll inside that building. The Lares are easily bribed and easily offended; neglecting to offer after a success ticks the Watch Patrol +1.
\item \textbf{The Ghosts of the Ninth:} They do not grant power, but they remember oaths. A character who swears a vow on the gate's keystone (a true vow, not a lie) hears a whisper: \textit{``We witness.''} Breaking the vow inflicts 2 Fatigue and the ghost's curse ( --1 die on all rolls until the vow is fulfilled).
\end{itemize}

\section*{GM Guidance: Magic in Intrigue Adventures}

Magic can short-circuit intrigue --- a simple \textit{Zone of Truth} or \textit{Invisibility} can bypass hours of negotiation. But intrigue is not about obstacles; it is about consequences. Let magic succeed, but let its use create new problems.

\begin{itemize}
\item \textbf{Make magic visible to those who look.} Cassia's silver eye sees arcane residue. A guard who falls asleep from a Cantor's song may later describe the feeling to a Censor's clerk. The party should never be sure they were not observed.
\item \textbf{Debt is the cost of convenience.} A Rite that opens a lock is faster than picking it. But the Traveler remembers. Mark Obligation, and let that Obligation come due at the worst moment.
\item \textbf{The curse is a mirror.} Any magic that touches the Iron Seal reflects back on the caster. The GM may offer a Devil's Bargain: \textit{``You can open the chest automatically --- but the seal's curse will affect you instead of the thief. Do you accept?''}
\item \textbf{Let magic fail interestingly.} A missed roll on a Livaean Rite does not mean ``nothing happens.'' It means the target is flattered but not fooled, or they recognize the enchantment and now treat the caster as a threat. The timer advances; the fiction deepens.
\end{itemize}

\section*{A Final Word}

Mareth's letter ends with a postscript, written in a different ink --- paler, as if mixed with ash.

\begin{quote}\itshape
I lied when I said I would not come. I am watching. I am always watching. The gate is a threshold, and thresholds are where I am strongest. When you hear the ninth bell, do not turn around. I will be there.

Do not be afraid. I am not the one you should fear.

\textit{The road continues.}
\end{quote}

The party may glimpse a hooded figure on the wall-walk, or find a small offering --- a knotted cord, a pinch of salt --- left at their camp. Mareth is never far from a crossroads.

\end{document}