Here’s a pdflatex-safe version (no fontspec, no \setmainfont, no \filcenter, and the bullets are converted to \textbullet):

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{sectsty}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{array}

\sectionfont{\Large\bfseries}
\subsectionfont{\bfseries}

\titleformat{\section}{\large\bfseries\centering}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\itshape}{}{0em}{}

\setlist{itemsep=0pt}
\setlist[enumerate]{itemsep=0pt}

\newcommand{\divider}{\vspace{12pt}\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}\vspace{12pt}}

\begin{document}

\begin{center}
{\LARGE \textbf{The Carnival of Broken Dreams}} \\
\vspace{6pt}
{\large A Folk Horror \& Social Commentary adventure for Fate's Edge} \\
\vspace{6pt}
{\large Tiers: I--III \textbullet{} Length: 3--6 sessions \textbullet{} Themes: Temptation, public goods, predatory spectacle}
\end{center}

\divider

\subsection*{What this is}

A traveling night-market of tents, music, and miracles glides the trade-roads between Silkstrand and Thepyrgos. It grants wishes---precisely---and charges a price in things people didn't know they valued. As the Desire Index rises, a town's intangibles unravel: lullabies vanish, bell-rights lapse, markets drift off schedule. The party can expose terms, retune the engine, or divert its appetite before the carnival moves on.

This module leans on core Fate's Edge procedures:
\begin{itemize}
\item Position \& DV: earn leverage first (permits, bells, seats), then roll.
\item Clocks: visible pressure on community, appetite, and exposure.
\item SB spends (on 1s): crowds shift, contracts ``correct,'' law turns mid-scene.
\item Diamonds: codified outcomes (e.g., Witness Seal, Stage Permit) that move Position without a roll.
\end{itemize}

\divider

\subsection*{Where it fits (Setting)}
\begin{itemize}
\item Silkstrand bridge-wards crave spectacle and short-cuts to prestige; Diamonds like Bellmark, Stage Permit, and Witness Seal reflect local bureaucracy and show-rights.
\item Thepyrgos law-by-bell and stair-rights make crowd scenes volatile; a single peal can flip Position or freeze a performance.
\item Patron Orbit (Lore): The carnival often courts Livaea (Desire) devotees and accepts etiquette enforcement by Mab's envoys---neither must be the villain. Treat them as cultural gravity and flavor, not rules engines.
\end{itemize}

\divider

\subsection*{Adventure Summary (at a glance)}

\textbf{Act I --- Firelight} \\
Parade arrives; a ``free wish'' hooks an important NPC. The party sees a miracle, then the missing price.

\textbf{Act II --- Mirrors \& Ledgers} \\
Investigate tents, negotiate terms, steal leverage from the Ledger Wagon, and win Diamonds to constrain predation.

\textbf{Act III --- The Calliope Heart} \\
A public performance powers the engine. Break one pillar---Rhythm, Mask Index, or Ticket Chain---to starve it or retool it.

\divider

\subsection*{Core Clocks (start values)}
\begin{itemize}
\item \textbf{Desire Index [6]} -- local wish-rate and price magnitude.
\item \textbf{Community Unraveling [6]} -- customs, names, and bonds fray.
\item \textbf{Carnival Appetite [8]} -- grand act when filled.
\item \textbf{Spotlight Heat [4]} -- how targeted the party is by staff.
\end{itemize}

\textbf{Tick examples}
\begin{itemize}
\item Desire Index: each granted wish (+1), heavy crowd teaser (+1 if unopposed).
\item Unraveling: any wish taken without informed terms (+1), public goods priced (+1).
\item Appetite: each botched sabotage the engine can ``eat'' (+1), retuning misfire (+1).
\item Heat: public confrontations, breaking taboos, or humiliating staff (+1).
\end{itemize}

\divider

\subsection*{The Desire Tax (custom procedure)}

\textbf{Idea:} Every fulfilled wish costs something the wisher didn't know they valued. As more wishes accrue, the Desire Index rises and prices escalate.

\textbf{Per Wish (run this)}
\begin{enumerate}
\item State the desire. Tag: Need, Ambition, or Indulgence.
\item Set the base price. Choose/roll on Taxable Intangibles (below).
\item Apply rate. Tax Magnitude = 1 + Desire Index. Scale the loss accordingly.
\item Resolve. The carnival fulfills; the price disappears from the wisher or town.
\item Advance clocks. Desire Index +1; Unraveling +1 if terms weren't fully understood; Appetite +1 on greedy staff actions.
\end{enumerate}

\textbf{Taxable Intangibles (d10)}
\begin{enumerate}
\item Private memory
\item Unconscious skill
\item Social grace
\item Local luck/wind
\item Shared custom
\item Legal right (bellmark/stair-pass)
\item Relationship texture
\item Place-name
\item Future possibility
\item Invisible boundary (e.g., where a river ``should'' bend)
\end{enumerate}

\textbf{Fallout thresholds}
\begin{itemize}
\item 0--1: Rumors, coincidences, sampler wishes.
\item 2--3: Faces feel unfamiliar; crowds drift to the midway.
\item 4--5: Names slip; permits misread; bells mistime.
\item 6+: Foundational loss unless stopped.
\end{itemize}

\textbf{Countermeasures}
\begin{itemize}
\item Informed consent: If the exact price is spoken and voluntarily accepted $\rightarrow$ reduce Magnitude by 1 (min 1) and don't tick Unraveling.
\item Substitution: PCs may volunteer an equal-value price to block Appetite.
\item Audited wish: Present a Witness Seal (Diamond) to force staff to state the exact terms; --1 DV vs carnival staff.
\end{itemize}

\divider

\subsection*{Diamonds (codified leverage)}

\begin{tabular}{|l|l|l|}
\hline
\textbf{Diamond} & \textbf{What it does} & \textbf{Typical Position shift} \\
\hline
Witness Seal & Compels staff to name the exact price aloud. & Controlled $\rightarrow$ Dominant (social) \\
Stage Permit & One lawful interruption of a live act. & Controlled $\rightarrow$ Dominant (on-stage) \\
Bellmark of Honest Trade & Shields from predatory refinances. & --1 DV for term disputes \\
Mirror Warrant & Break a mirror without accruing Heat. & Ignore one glamour penalty \\
Calliope Key & Retune the engine's rhythm for a beat. & Group +1 Position vs engine hazards \\
\hline
\end{tabular}

Hand out Diamonds as scene goals or rewards for clever prep. They change Position first; then roll.

\divider

\subsection*{Sites \& set pieces}

\textbf{Firelight Parade (opening)}

Mood: Lanterns, brass, thrilled hush. \\
Default Position: Controlled (DV 3). Public accusations raise DV to 4. \\
Beats: Free raffle-wish ``won'' by a sympathetic NPC; staff ``clarifies'' nothing. \\
SB spends (on 1s): Confetti seals a micro-contract; crowd tilts; a bell cue flips the scene order.

\textbf{Midway of Masks}

Try-before-buy illusions ``prove'' the wish.
\begin{itemize}
\item Risk: Each test without refusal ticks Desire Index +1.
\item Fail-forward: Small boon gained but a habit is lost (whistling, prayer, smile cadence) until resolved.
\end{itemize}

\textbf{House of Honest Mirrors}

Shows visitors as they believe themselves to be; mirrors hold their price ledger.
\begin{itemize}
\item Position: Controlled (DV 4).
\item Action: Break a mirror under Mirror Warrant to read its ledger safely.
\item Heat on miss: Mirror shards animate and reassign masks mid-scene.
\end{itemize}

\textbf{The Wishwright's Tent}

Velvet-law negotiation.
\begin{itemize}
\item Leverage: Witness Seal or Bellmark shifts Position; otherwise staff ``assumes the difference.''
\item Magic: Divination finds missing clauses but attracts Livaea-touched attendants.
\end{itemize}

\textbf{Ledger Wagon (backstage heist)}

A traveling records office; prices are written as negative space on vellum.
\begin{itemize}
\item Position: Controlled $\rightarrow$ Desperate if loud (DV 4--5).
\item Reward: Copying a ledger grants leverage to unwind a price; the reader temporarily loses a related social grace.
\end{itemize}

\textbf{The Calliope Heart (finale)}

Engine under the stage with three pillars:
\begin{itemize}
\item Rhythm: crowd cadence; break with counter-song (Cantor), bell off-beat, or Calliope Key.
\item Mask Index: who plays which role; scramble labels or sanctify a mask.
\item Ticket Chain: consent tokens; reroute or void them with contract sigils.
\end{itemize}

Break any one pillar to starve the engine or change its appetite.

\divider

\subsection*{NPC Roster}

\textbf{Maestra Pann (Ringmaster)} --- velvet coaxing, bell cues, spectacle first.
\begin{itemize}
\item Wants: Growth, clean shows, no scandal.
\item Moves: Change the beat; elevate stakes; flatter a rival into compliance.
\item Tell: Two-finger bell count before a hard turn.
\end{itemize}

\textbf{Mr. Vellum (Ledger-Clerk)} --- papery whisper, ink under nails.
\begin{itemize}
\item Wants: Balanced books (ethos of fairness... on their terms).
\item Moves: ``Correct'' a contract; split a cost between bystanders; misfile your name.
\item Tell: Checks quill nib before lying.
\end{itemize}

\textbf{The Cotton Witch (Livaea-touched)} --- sugared lips, cravings-as-magic.
\begin{itemize}
\item Wants: To feed your appetite until you consent to anything.
\item Moves: Offer a sweeter ``trial''; mirror your desire; name your envy aloud.
\item Tell: Hums when you hesitate.
\end{itemize}

\textbf{Harlequin of Courts (Mab's Envoy)} --- etiquette duels as combat.
\begin{itemize}
\item Wants: Order, pretty rules, public victories.
\item Moves: Demand apology; set a forfeit; bind you to a bit of theatre.
\item Tell: Bows before cuts.
\end{itemize}

\textbf{Ticket Twins (Ushers)} --- synchronized blinks, bright patter.
\begin{itemize}
\item Wants: Throughput, safety theatre.
\item Moves: Reassign seating; isolate meddlers; ``check'' tokens that edit consent after the fact.
\end{itemize}

Add one local anchor NPC (Bellwarden, Market Matron, or a child prodigy) to personify the stakes.

\divider

\subsection*{Procedures \& guidance}

\textbf{Position \& DV (quick ladder)}
\begin{itemize}
\item Dominant: Tier+2 --- permits in hand, crowd yours, law on your side.
\item Controlled: Tier+3 --- proper intros, stable ground.
\item Desperate: Tier+4+ --- taboo broken, wrong place/time, Heat high.
\end{itemize}

Use Diamonds to change Position before rolling. Reward prep with downshifts.

\textbf{SB spend menus (sample)}

\textbf{Parade Grounds}
\begin{itemize}
\item Free wish ``won'' by an allied NPC $\rightarrow$ Desire Index +1.
\item Crowd surge separates PC from their leverage.
\item Confetti-contract sticks to a sleeve: tear it = social harm.
\end{itemize}

\textbf{Backstage}
\begin{itemize}
\item Ledger page sticks to a shadow: lose a habit until returned.
\item A bell cue retimes the scene: +1 DV until the beat is broken.
\item Mask roles swap: allies briefly count as ``audience.''
\end{itemize}

\divider

\subsection*{Act structure (with beats)}

\textbf{Act I --- Firelight (1--2 sessions)}
\begin{itemize}
\item Hook via raffle-wish to a sympathetic townie.
\item Identify the Desire Tax through consequences.
\item Win first Diamond (Witness Seal) in a fair-game contest or etiquette duel.
\item Clocks: Desire Index 1--2; Unraveling 1; Heat 1.
\end{itemize}

\textbf{Fail-forward examples}
\begin{itemize}
\item Calling out the price in public: you force a lesser cost to be named, but it latches onto a bystander unless countered.
\item Breaking the line: you reach the tent; a child slips toward a booth you shut down.
\end{itemize}

\textbf{Act II --- Mirrors \& Ledgers (1--3 sessions)}
\begin{itemize}
\item House of Honest Mirrors $\rightarrow$ Wishwright's Tent $\rightarrow$ Ledger Wagon heist.
\item Earn Stage Permit or Mirror Warrant.
\item Decide whether to regulate, starve, or divert the engine.
\item Clocks: Desire Index 3--4; Unraveling 3--4; Appetite 3--5; Heat 2--3.
\end{itemize}

\textbf{Fail-forward examples}
\begin{itemize}
\item Mirror smash: gain a ledger shard; your reflection walks off (lose a habit) until reunited.
\item Retune calliope: DV --1 but Appetite +1 (the engine takes a taste).
\end{itemize}

\textbf{Act III --- The Calliope Heart (1 session)}
\begin{itemize}
\item Public show as ritual; break Rhythm, Mask Index, or Ticket Chain.
\item Choose an ending and settle prices.
\end{itemize}

\divider

\subsection*{Endings}
\begin{itemize}
\item \textbf{Expose \& Regulate:} Carnival stays but all wishes require informed, spoken terms. Unraveling clears; Livaea's devotees take interest in you.
\item \textbf{Starve the Engine:} Break any pillar; tents collapse; debts become letters and must be hand-delivered back.
\item \textbf{Divert the Appetite:} Offer a greater, voluntary group vow; the carnival leaves on a promise and returns when it matures.
\end{itemize}

\divider

\subsection*{Magic \& Talents (table-facing prompts)}
\begin{itemize}
\item Cantors: Lead counter-cadence; --1 DV vs crowd sway. Risk: echo-choir ticks Desire Index.
\item Runekeepers: Stamp fair-contract sigils; negate one clause (Obligation may rise if Courts are offended).
\item Invokers: Call a neutral market spirit to enforce equivalence; it tithes your luck.
\item Casters: Euclidean wards block mirror tricks; wards hum appetites that draw the Cotton Witch.
\item Summoners: Bind a ``lost-and-found'' spirit to return a price; it pockets a token from you.
\end{itemize}

\divider

\subsection*{XP milestones (suggested)}
\begin{itemize}
\item Name a Price (publicly, precisely): +6 XP
\item Save a Wisher (undo a price without harming others): +8 XP
\item Backstage Leverage (copy/steal ledger): +6 XP
\item Retune the Engine (break a pillar or reduce Desire Index to 0): +10 XP
\item Community Heals (clear Unraveling before departure): +12 XP
\item Session Base: +6 XP per session for exploration, roleplay, and setbacks
\end{itemize}

\divider

\subsection*{Random tables}

\textbf{d6 Midway Booths}
\begin{enumerate}
\item Guess-the-Secret (mirror reads you)
\item Cotton Alchemy (flavors as feelings)
\item Fortune Stair (mini-stair-right lottery)
\item Prize Hook (catch your ``best self'')
\item Lantern Maze (lefts are regrets)
\item Mask Painter (face becomes contract text)
\end{enumerate}

\textbf{d6 Crowd Moments}
\begin{enumerate}
\item Chant swells off-beat
\item Bell rings early
\item Child returns with a stranger's voice
\item Vendors swap prices mid-sentence
\item Duel of apologies erupts
\item Parade pauses; all eyes on you
\end{enumerate}

\textbf{d6 Prices That Hurt Just Right}
\begin{enumerate}
\item The last note of a cradle-song
\item Market wind that smells of dye
\item The shortcut only you knew
\item Your knack for remembering faces
\item The right to ring the bell once
\item The name of the street you grew up on
\end{enumerate}

\divider

\subsection*{Session 1 quickstart (3 hours)}
\begin{enumerate}
\item Cold Open (15 min): Parade arrives; raffle-wish triggers.
\item Investigation Beat (45 min): Midway of Masks or Honest Mirrors.
\item Pressure Beat (30 min): Public call-out; win Witness Seal.
\item Choice Beat (30 min): Who to help first; start or cool a clock.
\item Climax (30--40 min): Mini-debate or mirror heist.
\item Fallout (10 min): Hand out a Diamond; foreshadow the Calliope.
\end{enumerate}

\divider

\subsection*{Safety \& tone}
\begin{itemize}
\item This arc spotlights exploitation and ``price of desire.'' Use upfront consent tools. Swap any personal-price beats for harmless-but-meaningful community intangibles (songs, routes, customs) when that serves your table.
\end{itemize}

\divider

\subsection*{Printables (drop-in)}

\textbf{Clocks}
\begin{itemize}
\item Desire Index [\,] [\,] [\,] [\,] [\,] [\,]
\item Community Unraveling [\,] [\,] [\,] [\,] [\,] [\,]
\item Carnival Appetite [\,] [\,] [\,] [\,] [\,] [\,] [\,] [\,]
\item Spotlight Heat [\,] [\,] [\,] [\,]
\end{itemize}

\textbf{Diamonds (handouts)}
\begin{itemize}
\item Witness Seal \quad Stage Permit \quad Bellmark of Honest Trade \quad Mirror Warrant \quad Calliope Key
\end{itemize}

\divider

\subsection*{Ties \& follow-ons}
\begin{itemize}
\item Silk \& Velvet Arc: Bell-law and etiquette duels at the carnival foreshadow Crimson Masque.
\item Steps Into Sorrow: If a town loses its lullaby or street-name, those shards can seed sorrow-echoes on the Bridge or in Seven-Sorrows.
\end{itemize}

\divider

\subsection*{Running note}

Keep rolls scarce by rewarding preparation with Position shifts. Spend 1s to move the crowd, ``correct'' papers, or hit a new bell cue that alters scene order. The carnival is a machine for desire---let the party decide whether to smash, regulate, or re-aim it.

\end{document}