```latex
\documentclass[11pt,twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{fancyhdr}

\geometry{
    a4paper,
    left=25mm,
    right=25mm,
    top=25mm,
    bottom=25mm
}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={The Carnival of Echoes},
    pdfauthor={Fate's Edge Adventure}
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{The Carnival of Echoes}
\fancyhead[RE,LO]{A Fate's Edge Adventure}
\fancyfoot[C]{\thepage}

\titleformat{\section}
{\Large\bfseries}
{\thesection}
{1em}
{}

\titleformat{\subsection}
{\bfseries}
{\thesubsection}
{1em}
{}

\setlist{itemsep=0.5em}

\title{\textbf{The Carnival of Echoes}\\A Fate's Edge Adventure for Tiers I--II}
\author{}
\date{}

\begin{document}

\maketitle

\begin{center}
\textbf{A Stepwise Adventure Design}
\end{center}

\section*{STEP 1 --- Define the Core Premise}

\subsection*{1. What is happening?}
A traveling carnival that grants wishes is stealing memories and identities from visitors, leaving hollow shells behind.

\subsection*{2. Why now?}
The carnival arrives in town tonight for a three-day festival, and its malevolent attraction grows stronger with each victim---it plans to consume the entire town's collective memories by the final night.

\subsection*{3. Why the PCs?}
Each PC received a mysterious invitation to the carnival, and one of them has been having prophetic dreams about it. Local authorities are mysteriously absent or unresponsive.

\section*{STEP 2 --- Establish Setting \& Tone}

\subsection*{4. Where does the adventure occur?}
The riverside town of Millhaven during its annual Harvest Festival:
\begin{itemize}
    \item \textbf{Smell:} Cinnamon and roasting chestnuts mixed with something metallic
    \item \textbf{Sound:} Calliope music that sounds slightly off-key, distant carnival barker's voice
    \item \textbf{Architecture:} Victorian-era buildings with paper lanterns, the carnival's black-and-gold tents visible on the outskirts
    \item \textbf{Culture:} Close-knit farming community with superstitions about ``traveling folk''
    \item \textbf{Sight:} Strange shadows that don't match their owners in the evening light
\end{itemize}

\subsection*{5. What is the tone/theme?}
Psychological horror with elements of wonder and melancholy. Themes of identity, memory, and the price of desire.

\subsection*{6. What environmental pressures exist?}
\begin{itemize}
    \item The carnival only operates at night
    \item Town curfew enforced by mysterious carnival ``security''
    \item Increasing difficulty distinguishing real memories from false ones
    \item Growing paranoia as townspeople become ``hollowed''
\end{itemize}

\section*{STEP 3 --- Build Key NPCs}

\subsection*{Quest-giver: Mayor Elena Vasquez}
\begin{itemize}
    \item \textbf{Goal:} Save her town and her missing daughter
    \item \textbf{Fear:} Losing her remaining family and reputation
    \item \textbf{Method:} Gathering evidence while maintaining public order
    \item \textbf{Secret:} She made a wish 20 years ago that started this curse
    \item \textbf{Connection:} Knows one PC's family or has information they need
\end{itemize}

\subsection*{Local ally: ``Granny'' Moira Thornfield}
\begin{itemize}
    \item \textbf{Goal:} Protect the town's children and uncover the truth
    \item \textbf{Fear:} Becoming like her ``hollowed'' neighbors
    \item \textbf{Method:} Folk magic and local knowledge
    \item \textbf{Secret:} She's 153 years old and has seen this carnival before
    \item \textbf{Connection:} Offers sanctuary and information to the PCs
\end{itemize}

\subsection*{Rival group: The Hollowed Townsfolk}
\begin{itemize}
    \item \textbf{Goal:} Protect the carnival (they can't remember why they want to)
    \item \textbf{Fear:} The return of their painful memories
    \item \textbf{Method:} Subtle sabotage and misinformation
    \item \textbf{Secret:} They're aware they're compromised but can't resist the carnival's pull
    \item \textbf{Connection:} Former friends, family, or contacts of the PCs
\end{itemize}

\subsection*{Villain: The Carnival Master (Mr. Dram)}
\begin{itemize}
    \item \textbf{Goal:} Collect enough memories to become a god of forgotten dreams
    \item \textbf{Fear:} Being forgotten himself
    \item \textbf{Method:} Manipulation through desire and fear
    \item \textbf{Secret:} He's a construct made from the collective forgotten memories of previous victims
    \item \textbf{Connection:} Knows each PC's deepest desires and fears
\end{itemize}

\subsection*{Lieutenants:}
\begin{enumerate}
    \item \textbf{The Fortune Teller (Madame Zara)} - Reads victims' desires to tailor the perfect wish trap
    \item \textbf{The Strongman (Bull) Carson} - Enforces the carnival's will through intimidation and violence
    \item \textbf{The Carousel Operator (Twins)} - Literally spins victims' memories away through hypnotic rides
\end{enumerate}

\section*{STEP 4 --- Define the Villain's Plan}

\subsection*{7. What does the villain want?}
To collect 1,000 souls' worth of memories to ascend to godhood, becoming the eternal keeper of forgotten dreams.

\subsection*{8. Why can't they be stopped easily?}
The carnival exists partially outside normal reality and feeds on the desires that draw people to it. Destroying it requires breaking the collective belief in its magic.

\subsection*{9. What are the phases of their plan?}

\textbf{Setup (Night 1):} Carnival arrives, begins subtle memory collection through ``free'' wishes. Victims seem happier but lose small memories.

\textbf{Escalation (Night 2):} More aggressive collection, ``hollowed'' townspeople begin to act strangely. The carnival expands, offering more tempting wishes.

\textbf{Crisis (Night 3):} Final ritual to consume all remaining memories in town. Hollowed townsfolk become actively hostile to interference.

\section*{STEP 5 --- Map Out Player Choices}

\subsection*{10. What are the first three meaningful choices?}
\begin{itemize}
    \item Investigate the carnival's arrival and strange effects
    \item Confront a hollowed friend/family member for information
    \item Accept a tempting wish from the carnival to gain an advantage
\end{itemize}

\subsection*{11. What are the secondary choices?}
\begin{itemize}
    \item Rescue specific townspeople before they're fully hollowed
    \item Choose which memories to sacrifice to gain carnival knowledge
    \item Decide whether to destroy the carnival or try to reform it
\end{itemize}

\subsection*{12. What consequences do choices cause?}
\begin{itemize}
    \item Accepting wishes makes PCs more vulnerable to memory manipulation
    \item Rescuing townspeople gains allies but alerts the carnival
    \item Destroying the carnival without understanding it may leave fragments that return
\end{itemize}

\section*{STEP 6 --- Design the Core Mystery}

\subsection*{13. What is the central unknown?}
How the carnival steals memories and what happens to them, plus the Carnival Master's true identity and weakness.

\subsection*{14. What clues lead to answers?}
\begin{itemize}
    \item Hollowed townspeople speak in fragmented sentences about ``beautiful dreams''
    \item Granny Moira's old journals mention a similar carnival from decades ago
    \item The carnival's financial records (if accessed) show impossible payment methods
\end{itemize}

\subsection*{15. How do the players uncover the truth?}

\textbf{Three Clue Rule - The Carnival's True Nature:}
\begin{enumerate}
    \item \textbf{Investigation:} Research the carnival's past appearances reveals pattern of memory-theft
    \item \textbf{Social:} A partially hollowed NPC reveals the carnival feeds on forgotten dreams
    \item \textbf{Occult:} Magical investigation shows the carnival exists partially in the realm of dreams
\end{enumerate}

\section*{STEP 7 --- Build Encounter Variety}

\subsection*{16. Tactical Combat Encounter}
Battle with Bull Carson and hollowed townsfolk in the Strongman tent---fighting while resisting memory manipulation effects.

\subsection*{17. Exploration/Traversal Scene}
Navigating the ever-shifting carnival grounds where paths loop and familiar landmarks become threatening.

\subsection*{18. Social negotiation or deception}
Negotiating with Madame Zara for information while she tries to trick them into making wishes.

\subsection*{19. Set-piece or unique challenge}
Racing against time on the Carousel of Echoes---must escape before it ``spins away'' their memories permanently.

\subsection*{20. Skill challenge or hazard}
Disabling the carnival's dream-weaving machinery in the back areas while avoiding supernatural security measures.

\section*{STEP 8 --- Clarify the Stakes}

\subsection*{21. What happens if players succeed?}
The carnival is destroyed or reformed, memories are restored, but the town is left with trauma and some permanent changes. PCs gain reputation as protectors.

\subsection*{22. What happens if players fail?}
The carnival consumes the town's memories, turning everyone into hollow shells. It moves on to the next location with increased power.

\subsection*{23. What happens if players walk away?}
The carnival succeeds in Millhaven and grows stronger, becoming a recurring threat in future adventures. The PCs' hometown connections are lost.

\section*{STEP 9 --- Add a Midpoint Twist}

\subsection*{24. What ``oh shit'' moment flips assumptions?}
Halfway through Night 2, the PCs discover that one of their own has been making wishes at the carnival, and their memories are already compromised. They must decide whether to trust this PC or consider them compromised.

\section*{STEP 10 --- Prepare Player Spotlights}

\subsection*{25. Where is their personal moment?}

\begin{itemize}
    \item \textbf{Investigator/Tracker PC:} Gets to use their skills to uncover the carnival's pattern of previous victims across different towns
    \item \textbf{Magic User PC:} Faces a moral dilemma about using carnival magic to fight carnival magic
    \item \textbf{Social/Charismatic PC:} Must convince a hollowed loved one to remember who they really are
    \item \textbf{Combat Specialist PC:} Gets to fight the supernatural threat directly while protecting civilians
    \item \textbf{Stealth/Skill PC:} Infiltrates the carnival's backstage areas to disable its memory-stealing machinery
\end{itemize}

\section*{STEP 11 --- Final Confrontation + Dilemma}

\subsection*{26. What is the final set-piece?}
The Dream Consumption Ritual in the carnival's central tent---a swirling vortex of stolen memories where the Carnival Master grows more powerful as more townspeople are drawn in.

\subsection*{27. What choice defines the finale?}
To destroy the carnival completely (saving the town but losing the power it offered) or make one final wish to become the carnival's new masters (gaining great power but at an unknown cost to their humanity).

\section*{STEP 12 --- Aftermath \& Ripple Effects}

\subsection*{28. Who changes?}
\begin{itemize}
    \item The PCs gain or lose memories based on their choices
    \item Hollowed townspeople slowly recover (or don't, becoming NPCs)
    \item Granny Moira reveals more of her true nature and age
    \item Mayor Vasquez must deal with the political fallout
\end{itemize}

\subsection*{29. What alliances form or break?}
\begin{itemize}
    \item Strong bond with Granny Moira if she survives
    \item Tension with townspeople who made wishes they now regret
    \item Potential rivalry with other monster hunters who arrive too late
    \item New understanding of how supernatural threats work
\end{itemize}

\subsection*{30. What new adventure seeds emerge?}
\begin{itemize}
    \item Other carnival fragments survived and are appearing elsewhere
    \item Granny Moira's past with similar threats
    \item The Carnival Master's true origins in the realm of forgotten dreams
    \item PCs' own compromised memories may resurface later
\end{itemize}

\section*{🧩 Optional Modules}

\subsection*{Memory Fragments Track}
PCs can collect pieces of stolen memories as assets, but each fragment comes with emotional baggage.

\subsection*{Hollowing Clock}
[6] segments tracking how close the town is to total memory consumption.

\subsection*{Wish Corruption}
Each wish made at the carnival grants power but creates a corresponding weakness or moral compromise.

\subsection*{Dream Echoes}
PCs begin experiencing dreams of other victims' memories, gaining insight but risking their own identity.

\vspace{2em}

\begin{center}
\textbf{Total estimated play time: 3-4 sessions with room for expansion based on player choices and investigation depth.}
\end{center}

\end{document}
```