% The Crimson Ledger of Ecktoria - Fate's Edge One-Shot
% Author: [Your Name]
% License: CC-BY 4.0

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fontspec}
\usepackage{sectsty}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{tcolorbox}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{tikz}
\usepackage{array}
\usepackage{longtable}
\pagestyle{fancy}
\fancyhf{}
\rhead{Fate's Edge: The Crimson Ledger of Ecktoria}
\lhead{One-Shot Module}
\rfoot{\thepage}

\allsectionsfont{\centering\Large\bfseries}
\titleformat{\section}{\centering\Large\bfseries}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\bfseries}{}{0em}{}

\newtcolorbox{callout}[1][]{colback=blue!5!white,colframe=blue!75!black,fonttitle=\bfseries,title=#1}
\newtcolorbox{danger}[1][]{colback=red!5!white,colframe=red!75!black,fonttitle=\bfseries,title=#1}
\newtcolorbox{success}[1][]{colback=green!5!white,colframe=green!75!black,fonttitle=\bfseries,title=#1}
\newtcolorbox{patronbox}[1][]{colback=purple!5!white,colframe=purple!75!black,fonttitle=\bfseries,title=#1}

\begin{document}

\begin{center}
{\Huge \textbf{The Crimson Ledger of Ecktoria}} \\
{\Large A Fate's Edge One–Shot Adventure} \\
\textit{Political Intrigue in the Marble City}
\end{center}

\section*{The Pitch}
In the marble forums of Ecktoria, where the Utaran Empire's ghosts still echo through colonnaded halls, the city's greatest scandal isn't political—it's literary. The Crimson Codex, a legendary book of forbidden knowledge said to contain every secret, every lie, and every truth that could destroy the city's elite, has been stolen from the Library of Keys.

But this isn't just any theft. The Codex was bound in oaths—literally. Each page was a magically enforced secret, kept by the weight of sworn vows. Now those oaths are breaking, and as they do, Ecktoria's power structure is crumbling. Senators confess to treason in their sleep, merchants reveal the locations of their hidden vaults, and the High Priest of the Everflame begins preaching heretical sermons about the "true" history of the Empire.

\section*{The Hook}
The PCs are hired by Livaea, the Crimson Courtier herself (appearing as a patron to a desperate Runekeeper PC, or through a powerful NPC contact) to recover the Codex before the city tears itself apart. But Livaea has her own agenda—the Codex contains the location of the Red Branch, a splinter faction of the old Utaran nobility who vanished centuries ago but may hold the key to restoring the Empire.

The twist? The theft was orchestrated by Inaea, Angel of the Spider, working through a network of blackmail victims who finally found the courage to fight back. The breaking oaths aren't a bug—they're a feature. And the Codex is writing itself as it's read, with each secret revealed adding new pages of devastating truths.

\section*{What Makes It Cool}

\subsection*{Social Chaos as Adventure Hook}
Instead of monsters or armies, the threat is cascading social collapse. Noble houses turn on each other, merchant guilds expose each other's illegal dealings, and the city's famous bureaucracy begins processing absurd requests (like citizenship for statues, or marriage licenses for long-dead couples).

\subsection*{Living MacGuffin}
The Codex isn't just a book—it's actively rewriting itself, creating new secrets even as old ones are revealed. Players might find that reading certain passages changes their own character sheets or creates new complications.

\subsection*{Patron Theater}
Livaea and Inaea's conflict plays out through the PCs. Livaea wants order restored through manipulation and seduction of power, while Inaea wants the web of lies to collapse so new connections can form. Players must choose sides or forge their own path.

\subsection*{Genre Mashup}
High political intrigue meets magical mystery. Think "House of Cards" meets "The Name of the Wind" in a setting where words have weight and secrets are currency.

\subsection*{Ecktoria as Character}
The city itself becomes a reactive NPC. As oaths break, districts change personality—military quarters become paranoid, merchant districts turn anarchic, religious areas become fanatical.

\section*{Session Structure}

\subsection*{1. The Confessions}
Arrive to find the city's elite involuntarily spilling secrets, hired to investigate.

\subsection*{2. Threads of Power}
Follow the web of blackmail and counter-blackmail as alliances shift.

\subsection*{3. The Red Branch}
Discover the exiled nobles are trying to return through the chaos.

\subsection*{4. Patron's Gambit}
Livaea and Inaea make their final moves for control.

\subsection*{5. The New Ecktoria}
Decide what kind of city rises from the ashes of broken oaths.

\section*{Cool Set Pieces}

\begin{itemize}
    \item \textbf{A dinner party} where everyone reveals their darkest secrets in turn.
    \item \textbf{The Forum of Keys} where magical contracts are being nullified mid-signature.
    \item \textbf{A chase through the city's bureaucratic maze} as players try to stop a marriage between political rivals that would unite the city's factions.
    \item \textbf{The final confrontation in the Library} where the Codex is rewriting itself with each revelation.
\end{itemize}

\section*{Fronts \& Clocks}

\subsection*{Secrets Unbound [6/8]}
This clock tracks the cascading revelation of secrets throughout Ecktoria.

\textbf{Advances when:}
\begin{itemize}
    \item Major secret revealed by NPC or Codex reading
    \item Failed social manipulation roll (DV 4+)
    \item Patron Rite cast that affects social structures
    \item PCs choose Inaea's path over Livaea's path
\end{itemize}

\textbf{Segments:}
\begin{enumerate}
    \item Minor scandals emerge. Gossip spreads faster.
    \item Moderate revelations. One significant NPC makes public confession.
    \item Major social disruption. Alliance breaks or forms.
    \item City-wide chaos. Multiple factions in conflict.
    \item Institutional collapse. Key organizations cease functioning.
    \item Catastrophic breakdown. City government paralyzed.
\end{enumerate}

\subsection*{Codex Rewriting [4/6]}
This clock represents the Codex's active response to events.

\textbf{Advances when:}
\begin{itemize}
    \item PCs read 2+ pages from the Codex
    \item 3+ NPCs make major revelations in same scene
    \item Patron's agenda advances significantly
    \item Secrets Unbound clock advances 2+ segments
\end{itemize}

\textbf{Segments:}
\begin{enumerate}
    \item New minor secret appears in margins.
    \item Codex gains ability to reveal one specific truth.
    \item Major new chapter appears with game-changing information.
    \item Codex becomes actively hostile to readers.
\end{enumerate}

\section*{Key NPCs}

\subsection*{Livaea, the Crimson Courtier}
\textbf{Tags:} Patron, Social Manipulator, Elegant Destroyer \\
\textbf{Approach:} Presence 4, Wits 4, Spirit 3, Body 2 \\
\textbf{Skills:} Sway 5, Command 4, Deception 4, Lore 3 \\
\textbf{Talents:} Courtly Guise, Binding Vow, Sovereign Glamour \\
\textbf{Harm:} 6-clock; Armor: 2 \\
\textbf{Special:} Glamour's Touch - +1d to social rolls, but cannot speak plain falsehoods.

\textbf{Agenda:} Restore order through controlled revelation. Stabilize key alliances. Prevent complete collapse of Ecktorian institutions.

\subsection*{Inaea, Angel of the Spider}
\textbf{Tags:} Patron, Web-Weaver, Truth-Bringer \\
\textbf{Approach:} Wits 5, Spirit 4, Presence 3, Body 2 \\
\textbf{Skills:} Insight 5, Investigation 4, Sway 3, Lore 4 \\
\textbf{Talents:} Strand of Inevitability, Snaring Filament, Web-Sense \\
\textbf{Harm:} 5-clock; Armor: 1 \\
\textbf{Special:} Web-Sense - +2 dice to detect hidden connections and leverage.

\textbf{Agenda:} Expose all corruption. Break unjust bonds. Let the truth rebuild society from its foundations.

\subsection*{High Priest Valeria the Truth-Speaker}
\textbf{Tags:} Religious Authority, Heretic, Chosen Voice \\
\textbf{Approach:} Spirit 4, Presence 3, Wits 3, Body 1 \\
\textbf{Skills:} Command 4, Lore 5, Faith 4, Insight 3 \\
\textbf{Talents:} Divine Authority, Prophetic Sight, Sacred Texts \\
\textbf{Harm:} 3-clock; Armor: 0 \\
\textbf{Special:} Chosen Voice - Words carry divine weight, +2 Effect to declarations.

\section*{Patron Interaction}

\begin{patronbox}[Livaea's Offers]
\textbf{Binding Vow:} Stabilize a crumbling alliance. Gain +1d to social rolls with that faction, but mark 1 Obligation.

\textbf{Sovereign Glamour:} Create a perfect social situation. Next social scene begins with Dominant Position, but generate 1 SB (Hearts).

\textbf{Token of Favor:} Gain temporary ally's support. +1 Effect to one action, but the ally marks 1 Exposure.
\end{patronbox}

\begin{patronbox}[Inaea's Offers]
\textbf{Strand of Inevitability:} Reveal hidden connection. Gain +2 dice to Investigation about target, but mark 1 SB (Diamonds).

\textbf{Snaring Filament:} Set social trap. Next hostile NPC suffers -1d to social rolls, but you mark 1 SB (Spades).

\textbf{Merciful Severing:} Break harmful tie. Remove one Condition or SB, but mark 1 Fatigue from the release.
\end{patronbox}

\section*{Ecktoria Districts - Reactive States}

\subsection*{The Grand Forum (Law \& Order)}
\begin{tabular}{|p{2cm}|p{5cm}|p{5cm}|}
\hline
\textbf{State} & \textbf{Effects} & \textbf{Social DV Modifier} \\
\hline
Orderly & Bureaucracy functions normally & Base DV \\
Chaotic & Contracts failing, magistrates arguing & +1 DV \\
Purges & Zealous investigations, false accusations & Desperate Position \\
Reconciliation & Forgiveness sought, old debts forgiven & Controlled Position \\
\hline
\end{tabular}

\subsection*{The Coin-House Floor (Commerce)}
\begin{tabular}{|p{2cm}|p{5cm}|p{5cm}|}
\hline
\textbf{State} & \textbf{Effects} & \textbf{Financial DV Modifier} \\
\hline
Stable & Normal negotiations, sealed ledgers & Base DV \\
Anarchic & Frantic deals, bankruptcies & +2 DV \\
Speculative & Wild investments, rumors drive prices & -1 DV (but risk) \\
Collapse & Markets frozen, credit worthless & Desperate Position \\
\hline
\end{tabular}

\subsection*{The Basilica of the Everflame (Religion)}
\begin{tabular}{|p{2cm}|p{5cm}|p{5cm}|}
\hline
\textbf{State} & \textbf{Effects} & \textbf{Ritual DV Modifier} \\
\hline
Orthodox & Rigid doctrine, formal rites & Base DV \\
Ecstatic & Truth-seeking, spontaneous worship & +1 Effect \\
Doubting & Faith shaken, rituals uncertain & +1 DV \\
Fanatical & Zealous purges, forced conversions & Desperate Position \\
\hline
\end{tabular}

\section*{The Crimson Codex - Mechanics}

\subsection*{Reading the Codex}
\textbf{Risk:} Each reading requires Wits + Arcana DV 4. On failure, generate 1 SB. On partial, choose one:
\begin{itemize}
    \item Reveal truth but mark 1 Fatigue
    \item Gain insight but generate 1 SB (Diamonds)
    \item Learn secret but mark 1 SB (Hearts)
\end{itemize}

\textbf{Rewards:} Successful reading grants:
\begin{itemize}
    \item +1d to Investigation about related topics
    \item One specific truth about target NPC/faction
    \item Advantage on social rolls against target (once)
\end{itemize}

\subsection*{Codex Backlash Table}
When reading fails or generates SB, roll 1d6:
\begin{enumerate}
    \item Your secret revealed to nearby NPCs
    \item Target of your investigation becomes aware
    \item Patron (Livaea or Inaea) gains leverage over you
    \item New complication clock starts (2-4 segments)
    \item Harm 1 (Mental stress from forbidden knowledge)
    \item Gain permanent Condition: Truth-Sensitive (-1d to deception)
\end{enumerate}

\section*{Investigation and Social Challenges}

\subsection*{Following the Threads}
Use the Deck of Consequences for social interactions:
\begin{itemize}
    \item \textbf{Hearts:} Social/Emotional fallout - relationships strained, reputations damaged
    \item \textbf{Diamonds:} Magical/Spiritual disturbance - divine attention, supernatural interference
    \item \textbf{Clubs:} Material cost - financial losses, equipment damage
    \item \textbf{Spades:} Harm/Escalation - physical danger, situation worsening
\end{itemize}

\subsection*{The Web of Oaths}
Create a simple visual map of key NPCs and connections:
\begin{itemize}
    \item Political (red lines)
    \item Financial (gold lines)  
    \item Romantic (pink lines)
    \item Blackmail (black lines with arrows)
\end{itemize}

As oaths break, lines snap or change color. Players can:
\begin{itemize}
    \item \textbf{Sever} (Inaea): Cut one connection, but generate 1 SB
    \item \textbf{Reinforce} (Livaea): Strengthen one connection, but mark 1 Obligation
    \item \textbf{Redirect}: Change target of connection, but require Wits + Sway test
\end{itemize}

\section*{Potential Twists}

\subsection*{The Red Branch's True Agenda}
The exiled nobles aren't returning to restore the Empire—they want to reclaim their lost artifacts and disappear again, using the chaos as cover. The Codex might be their key to finding these items.

\textbf{Reveal:} In scene 3, when PCs investigate the Red Branch, they discover plans for a "quiet extraction" rather than a glorious return.

\subsection*{The Chosen Voice}
The High Priest isn't mad; she's been chosen by the Codex itself to be its voice, preaching the "truth" it contains.

\textbf{Reveal:} During a religious ceremony, her eyes turn crimson when reading from any text, and she speaks prophecies that later prove true.

\subsection*{Personal Secrets}
A PC's own secret is contained within the Codex. Reading it or being near it causes them to involuntarily reveal it, or grants the Patron (Livaea or Inaea) leverage over them.

\textbf{Implementation:} Each PC writes down one secret during character creation. When Codex Rewriting advances, GM may reveal one PC secret.

\subsection*{The Ultimate Trap}
The Codex is a trap, designed by an ancient enemy of the Utarans. The "secrets" are half-truths meant to lead Ecktoria (or its successor state) to its doom.

\textbf{Reveal:} Final confrontation reveals the Codex was created by the Sealed Gate to destabilize Utaran successor states through information warfare.

\section*{GM Guidance}

\subsection*{Running the City}
Make Ecktoria feel alive by:
\begin{itemize}
    \item Describing changing atmospheres in districts
    \item Having NPCs react differently based on district state
    \item Using the Secrets Unbound clock to trigger district shifts
    \item Letting player actions have visible consequences on city mood
\end{itemize}

\subsection*{Patron Balance}
Present both Patrons' offers as:
\begin{itemize}
    \item Equally tempting within their frameworks
    \item Logically consistent with their natures
    \item Carrying clear, immediate costs
    \item Leading to different long-term consequences
\end{itemize}

\subsection*{Social Combat}
Treat social encounters like combat:
\begin{itemize}
    \item Establish clear stakes and consequences
    \item Use Position to reflect social advantage
    \item Let clocks drive escalating tension
    \item Reward creative solutions and roleplay
\end{itemize}

\section*{Character Integration}

\subsection*{For Runekeepers}
The abundance of oaths and secrets makes Oath-based Rites incredibly potent:
\begin{itemize}
    \item \textbf{Binding Vow:} Stabilize crumbling alliances
    \item \textbf{Mark of Antlers:} Grant temporary authority in chaos
    \item \textbf{Revelation:} Pierce illusions but risk unwanted truths
\end{itemize}

\subsection*{For Cantors}
The Cantor's Path can be used to "sing" truths or lies:
\begin{itemize}
    \item \textbf{Hymn of Truth:} Force honesty in social encounters (+1 Effect)
    \item \textbf{Song of Lies:} Create convincing deceptions (+1d to Deception)
    \item \textbf{Chorus of Unity:} Calm conflicts (+1 Position to peacemaking)
\end{itemize}

\subsection*{For Invokers}
Ritual magic excels at investigation and information gathering:
\begin{itemize}
    \item \textbf{Reveal Rites:} Expose hidden connections or concealed truths
    \item \textbf{Ward Rites:} Protect against unwanted revelations
    \item \textbf{Command Rites:} Compel truth from unwilling sources
\end{itemize}

\section*{Session Flow Recommendations}

\subsection*{Opening (30 minutes)}
\begin{itemize}
    \item Establish the chaos through NPC reactions
    \item Present the initial hook from Livaea or contact
    \item Start Secrets Unbound clock at 2 segments
    \item Introduce 2-3 key NPCs with immediate problems
\end{itemize}

\subsection*{Midgame (2 hours)}
\begin{itemize}
    \item Focus on 2-3 major investigation threads
    \item Advance clocks through player choices
    \item Introduce Patron offers and complications
    \item Build toward Red Branch revelation
\end{itemize}

\subsection*{Climax (1 hour)}
\begin{itemize}
    \item Library confrontation with Codex
    \item Patron final gambits
    \item Player choice determines outcome
    \item Resolve major NPC storylines
\end{itemize}

\subsection*{Resolution (30 minutes)}
\begin{itemize}
    \item Show consequences of player choices
    \item Award XP for engagement and roleplay
    \item Discuss potential campaign expansion
\end{itemize}

\begin{callout}[Final Words]
Ecktoria will never be the same. Whether it becomes a beacon of truth, a web of new connections, or a cautionary tale depends on the choices your players make. The marble halls will echo with new stories—some of redemption, others of revelation, all written in the ink of consequence.
\end{callout}

\end{document}
