% kahfagia_adventure.tex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{setspace}
\onehalfspacing
\usepackage{microtype}
\usepackage{ebgaramond}
\usepackage{fancyhdr}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

\title{The Gilded Chain}
\author{A Fate's Edge Adventure for Tier II--III}
\date{}

\begin{document}

\maketitle
\tableofcontents

\section{Introduction}

\subsection{Overview}
The PCs are officers or crew aboard the Kahfagian frigate *Vigilant* (or a ship of their name). They serve Captain Talia Taylor, a veteran of the cold war with Oshiira. The Admiralty orders the *Vigilant* to the colonial port of **Calinda** on the Dhaharan Brass Coast to quell a "tax rebellion". The rebellion is actually a protest against the Admiralty's ruthless grain levies. The real instigator is a corrupt oligarch, **Lord Admiral Castellan**, who wants to provoke a war to cover up his embezzlement.

The crew must navigate conflicting orders, spy on their own superiors, and ultimately decide: follow orders and massacre civilians, mutiny and defect to the rebels, or turn pirate and sail the monsoon winds.

\subsection{Tone}
- Naval thriller with moral weight.
- Inspired by the complexities of colonialism: Kahfagia itself is a creole nation (mixed Oshiiran and Vilikari) but still commits imperial acts.
- The enemy is not the rebels but the system that profits from exploitation.
- Player choice drives the ending.

\subsection{Content Warnings}
Colonial violence, execution of civilians, betrayal. Use X‑Card and lines/veils.

\subsection{Running Time}
2--3 sessions.

\section{Background for the GM}

\subsection{Kahfagia: A Creole Empire}
Kahfagia was once an Oshiiran colony, settled by Vilikari merchants and Oshiiran administrators. Over generations, a mixed culture emerged -- speaking a creole language, following a blend of lantern‑law and ancestor rites. Its people are fiercely independent and proud of their "free" status. Yet they have built their own empire on the Dhaharan coast, extracting grain and labor. This irony is not lost on the rebels: "The freed enslaved the free."

\subsection{The Admiralty}
The Thalassocracy is ruled by the **Admiral of Admirals** (a figurehead) and a council of oligarchs -- merchant families who own the navy. The most powerful are Houses **Merino** (old money, legalistic) and **Velasquez** (new money, ruthless). They vie for control of the spice trade. Lord Admiral Castellan is a Velasquez crony.

\subsection{The Rebellion}
The port of Calinda is a Sidhi settlement forced to pay tribute in grain. When Admiralty inspectors demanded double last year, the Sidhi refused. Castellan declared them "tax rebels" and asked for military support. The *Vigilant* is the first ship to respond.

\subsection{The Truth Castellan Hides}
Castellan embezzled the grain that was supposed to feed the Admiralty's own fleet. He needs a war to blame the loss on pirates and tighten his grip.

\section{Player Characters (Pre‑gens or Created)}

Provide pre‑gens or allow players to make Tier II--III characters with Kahfagian backgrounds.

\begin{itemize}
    \item \textbf{Captain Talia Taylor (NPC)} -- Honourable, patriotic, torn.
    \item \textbf{First Mate Sebastian} -- Pragmatic, loyal to the crew.
    \item \textbf{Navigator Sora} -- Sidhi, knows the port, secretly connected to rebels.
    \item \textbf{Gunnery Officer Rigg} -- Scarred, hates oligarchs.
    \item \textbf{Surgeon Miri} -- Idealistic, morally conflicted.
\end{itemize}

\section{Adventure Structure}

\subsection{Act I: The Quiet Port}
The *Vigilant* docks at Calinda. The crew finds an eerie silence: no trade, no fishing boats. A delegation of Sidhi elders offers to pay the original tax if the Admiralty withdraws the double levy. Captain Taylor is sympathetic but bound by orders to enforce the increase.

PCs can:
- Investigate the town -- find a hidden shrine to Ikasha where rebels meet.
- Talk to rebels -- they have proof of Castellan's embezzlement.
- Spy on the Admiralty agent in town -- a nervous clerk with a sealed safe.

**Timers:**
- \textbf{Rebellion Ignites} [6] -- ticks if PCs harm civilians or ignore pleas.
- \textbf{Loyalty to the Captain} [6] -- ticks if PCs follow orders that compromise their morals.

**Key Scene:** The rebels offer the PCs a copy of Castellan's secret ledger in exchange for safe passage out of the port. If the PCs accept, they gain a powerful blackmail tool.

\subsection{Act II: The Admiral's Gambit}
A message from Castellan arrives via pigeon: "Execute the ringleaders at dawn." Captain Taylor is ordered to hang the rebel elders. She stalls, asking the PCs for advice.

PCs can:
- Confront the Admiralty agent -- he has the execution order, but he is also scared.
- Search the agent's safe -- requires a heist; contains letters proving Castellan's fraud.
- Sabotage the gallows -- delay execution until proof can be sent to the Admiral of Admirals.

**Twist:** The Admiralty agent is a Velasquez loyalist; he sends a runner to inform Castellan. The PCs must intercept or the runner will bring a warship.

**Timers:**
- \textbf{Castellan's Runner} [4] -- if filled, a frigate arrives (fight or surrender).
- \textbf{Proof to the Throne} [8] -- if filled, the Admiral of Admirals issues a recall of Castellan.

\subsection{Act III: The Choice}
Captain Taylor calls a ship's council. She reveals she has the proof. But the *Vigilant* is low on supplies, and the crew is divided.

Possible endings:

\begin{enumerate}
    \item \textbf{Follow Orders (Grim End)} -- Execute the rebels. The *Vigilant* becomes a symbol of terror. The crew loses morale; many desert. Captain Taylor resigns in shame.
    \item \textbf{Defect to the Rebels} -- The crew turns its guns on the Admiralty agent and declares its own flag. The *Vigilant* becomes a pirate hunter of the corrupt, hunted by the navy.
    \item \textbf{Trade on the Monsoon} -- The crew sells the proof to House Merino in exchange for a letter of marque. They become privateers for the rival oligarch, sailing south to Ayokha. A bittersweet freedom.
    \item \textbf{True Mutiny} -- The PCs lead a mutiny, arrest Captain Taylor (she agrees), and sail to the Oshiiran border to seek asylum. They become spies for the Widow.
    \item \textbf{The Noble Lie} -- They forge documents that the rebels paid the levy, then slip away. Calinda returns to silent misery, but no blood is shed. The crew is haunted.
\end{enumerate}

\subsection{Epilogue}
The GM narrates consequences based on the choice. The crew's reputation changes (Pirate, Traitor, Hero). Captain Taylor's fate is decided. The Kalrantha amulet is not involved, but a rumour of it may be seeded for a future campaign.

\section{NPC Gallery}

\begin{itemize}
    \item \textbf{Captain Talia Taylor} -- Middle‑aged, scarred, tired. Wants to serve Kahfagia, but not at the cost of her soul.
    \item \textbf{Lord Admiral Castellan} -- Never appears; his presence is felt through orders and letters. A voice of oily confidence.
    \item \textbf{Sister Tamsin} -- Rebel leader, a Sidhi priestess of Ikasha. She offers the ledger. Knows she may die.
    \item \textbf{Clerk Vel} -- Admiralty agent, young, terrified. Wants to go home. Can be turned.
    \item \textbf{First Mate Sebastian} -- Practical, stoic. His loyalty is to the crew, not the flag.
\end{itemize}

\section{Running the Adventure}

\subsection{Pacing Suggestions}
- Act I: investigation and social encounters (2 hours).
- Act II: heist and moral dilemmas (2 hours).
- Act III: council and choice (1 hour).

\subsection{Handling Player Agency}
The adventure is designed for moral weight. Encourage debate. Allow a flashback to a rebellious ancestor or a memory of Admiralty brutality. Rewards and consequences are emotional, not just mechanical.

\subsection{Alternate Hook}
If the party are not Kahfagian, they could be mercenaries hired by the Admiralty. Their loyalty is purely financial, which adds a different pressure.

\subsection{Cold War Context}
Mention the Oshiiran spies in Calinda (they are neutral in this affair; they only watch). A subplot could involve selling the proof to the Widow instead of the Admiral of Admirals.

\section{Rewards and Legacy}

\begin{itemize}
    \item If they defect or become privateers, they gain a ship tag: \textbf{Sea Wolf} (once per session, they may escape a naval encounter without rolling).
    \item If they turn pirate, they gain a Black Mark with Kahfagia (enemy fleets will pursue) but also a String: \textbf{Rebel Network} (safe harbour at Calinda).
    \item If they trade on the monsoon, they gain a Standard Asset: \textbf{Monsoon Contract} (a trading route that gives +1 Credit per leg).
\end{itemize}

\section{Conclusion}

\textit{The Gilded Chain} is a self‑contained moral thriller set against the backdrop of Kahfagia's creole empire. It asks: can a nation founded by the abused become an abuser itself? The players decide the answer---and live with the consequences.

\vspace{2\baselineskip}
\begin{center}
\textit{"The chain that bound you can be reforged. Or you can break it and sail away."}
\end{center}

\end{document}
