\documentclass[12pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{mdframed}
\usepackage{needspace}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{calc}
\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{longtable}
\usepackage{pdflscape}
\usepackage{multicol}
\usepackage{graphicx}

\definecolor{darkblue}{rgb}{0.0, 0.0, 0.55}
\hypersetup{
    colorlinks=true,
    linkcolor=darkblue,
    urlcolor=darkblue,
    pdftitle={The Gilded Heist},
    pdfauthor={Fate's Edge Adventure Module}
}

% ------------------- Custom commands -------------------
\newcommand{\tag}[1]{\textsc{#1}}
\newcommand{\clk}[1]{\textit{[#1]}}
\newcommand{\stringitem}[1]{\textit{#1}}
\newcommand{\dusana}[1]{%
    \begin{quote}\itshape #1 ---~Dusana of the Raven Road\end{quote}}
\newcommand{\letter}[2]{%
    \begin{quote}\small\itshape #1 \hfill --- #2\end{quote}}

% ------------------- Styling -------------------
\titleformat{\chapter}{\Huge\bfseries}{\thechapter}{20pt}{}
\titleformat{\section}{\Large\bfseries}{\thesection}{15pt}{}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8ex}

% ------------------- Infobox styling -------------------
\newmdenv[
    linecolor=black,
    backgroundcolor=gray!5,
    innertopmargin=8pt,
    innerbottommargin=8pt,
    innerrightmargin=8pt,
    innerleftmargin=8pt,
    skipabove=6pt,
    skipbelow=6pt,
    font=\small
]{infobox}

% ------------------- Document -------------------
\begin{document}

\title{
    \Huge\textbf{The Gilded Heist}\\[0.5em]
    \Large A Fate's Edge Adventure
}
\author{
    Tier II (2--3 Sessions)\\
    \small Silkstrand Intrigue Series
}
\date{
    \small Written on velvet paper
}
\maketitle

\tableofcontents

\chapter{Overview: The Economy of Favors}

Every lock has a key. Every guard has a price. And every clever plan leaves a thumbprint on someone's ledger. Livaea does not steal. She borrows---and the interest compounds in awkward dinner invitations.

\vspace{1em}
\dusana{I once watched a thief pick a lock with a hairpin. It took three seconds. I once watched a courtier pick a lock with a compliment. It took three years. Both openings lead to the same room. The difference is who you meet when you walk in.}

\section{Premise}

A notorious heist crew---that's the party---has been hired to steal the \textbf{Sunset Ledger} from the Amber Villa, a fortress-like manor in Silkstrand. The Ledger contains damning evidence of a merchant lord's tax evasion, but more importantly, it's sealed with a magical lock that requires a specific \textit{social key}: a secret handshake, a whispered compliment, a favor called in. The party's patron is a mysterious woman known only as \textbf{The Velvet Countess}---actually an Invoker who owes a deep debt to Livaea, the Crimson Courtier. Every clever plan the crew devises works flawlessly, but each success also creates a \textbf{social debt}: a favor owed to someone in Livaea's sprawling network. By the end, the party must decide whether to pay those debts, pass them on, or burn them---and risk Livaea's attention.

\section{Themes}

\begin{itemize}
    \item \textbf{Heist Tropes:} The plan, the montage, the double-cross.
    \item \textbf{Social Debt as Currency:} Favors are more valuable than gold.
    \item \textbf{The Invisible Economy:} Obligations bind people tighter than chains.
    \item \textbf{Parody:} A gentle mockery of \textit{gentleman thief} stories where the consequences are social comedy rather than gunfire.
\end{itemize}

\section{Tone}

Witty, tense, with a layer of ironic distance. The heist is played straight, but the consequences are social comedy. Think \textit{Ocean's 11} meets \textit{Dangerous Liaisons}. The danger is not death, but embarrassment---or worse, owing a dinner party to someone you dislike.

\section{Key Mechanics}

\begin{infobox}
\textbf{Adventure Mechanics}
\begin{itemize}
    \item \textbf{Social Debt Clock [6]:} Tracks the party's cumulative obligations. Each time they use a \textit{clever plan} that relies on a contact, a bribe, or a forged invitation, tick the clock. When full, Livaea's representative (or one of her debtors) calls in a favor---at the worst possible moment.
    \item \textbf{Favor Tokens:} Specific debts owed to named NPCs. Can be spent for automatic successes on related rolls, but each use ticks the Social Debt Clock again.
    \item \textbf{The Plan:} Borrowed from \textit{Blades in the Dark} flashback mechanics. Players can declare a preparation flashback by spending 1 Boon (or marking 1 Fatigue). But each flashback also creates a new Debt token with a random NPC.
\end{itemize}
\end{infobox}

\section{Locations}

\begin{itemize}
    \item \textbf{The Amber Villa:} Three floors: entry hall with dancing, offices with ledgers, vault in the basement.
    \item \textbf{The Velvet Countess's Townhouse:} Safe house, planning room, smells of jasmine and old paper.
    \item \textbf{The Crimson Court:} Livaea's social sphere, not visited but referenced constantly.
\end{itemize}

\section{Key NPCs}

\begin{itemize}
    \item \textbf{The Velvet Countess:} Livaea's agent, a Runekeeper with a Symbol of Livaea. Elegant, tired, and deeply in debt to her patron.
    \item \textbf{Lord Caspian Verre:} Target, a merchant prince with a secret: he's also an Invoker of Livaea, trading secrets for social power.
    \item \textbf{Mistress Thorn:} Head of security, a former Black Banner mercenary who loves riddles.
    \item \textbf{The Debt Collector:} A silent figure in a porcelain mask who appears whenever the Social Debt Clock fills; speaks only in questions.
\end{itemize}

\section{The Hook}

The party is summoned to the Velvet Countess's parlor. She has a job. The pay is excellent: 1,000 silver marks, plus a magical trinket of their choice from the vault. But she warns them: \textit{Livaea's favors are like silk gloves. They fit perfectly, but you forget you're wearing them until you can't feel your own fingers.}

\chapter{Act I: The Proposal}

\section{Scene 1: The Velvet Parlor}

The room smells of jasmine and old paper. The Velvet Countess pours tea from a pot that steams without a fire beneath it.

\letter{Lord Caspian Verre has offended someone who cannot be named. The Sunset Ledger, kept in his basement vault, details every bribe he has ever paid and every official he has ever compromised. We need it. Not for blackmail---for balance.}{--- The Velvet Countess}

She provides a dossier:
\begin{itemize}
    \item The Villa's layout (three floors, guards on rotation, a riddle-lock on the vault).
    \item The guest list for the annual Lantern Gala tomorrow night (the party's infiltration vector).
    \item A handful of \textbf{Favor Tokens} pre-issued: invitations to the gala, a guard's uniform, a forged letter of recommendation. Each token has a name attached (e.g., \textit{Favor owed to Lady Wren, who expects you at her salon next week}).
\end{itemize}

The party can spend these tokens immediately or save them. Using them ticks the \textbf{Social Debt Clock} by 1.

\section{Scene 2: The Plan (Montage)}

Players declare their approach (disguise, stealth, social engineering, or magical infiltration). The GM runs a quick flashback sequence where each player describes one \textit{clever preparation} (e.g., \textit{I bribed the chef to leave the kitchen door unlocked}). Each preparation costs 1 Boon or 1 Fatigue, and creates a new Favor Token with a random NPC (roll on the Favor Table). The GM notes the token and its implied cost.

\begin{infobox}
\textbf{Favor Table (d6)}
\begin{enumerate}
    \item A debt to a flirtatious dancer who now expects a date.
    \item A promise to sponsor a struggling artist's exhibition.
    \item An IOU to a guard who looked the other way---he wants the party's help escaping the city someday.
    \item A secret shared with a gossip columnist; she will print it unless the party gives her something juicier.
    \item A vial of expensive perfume \textit{borrowed} from a noble's wife; she will notice it missing.
    \item A favor owed to a priest of the Everflame, who now considers the party \textit{spiritually indebted}.
\end{enumerate}
\textit{The Social Debt Clock starts at 0 + number of tokens created.}
\end{infobox}

\section{Scene 3: The Gala (Infiltration)}

The party arrives at the Amber Villa. The party is in full swing; masked guests swirl. They must navigate three layers:

\begin{enumerate}
    \item \textbf{The Dance Floor (Social):} Gain access to the east wing by winning a dance contest, distracting a key guard, or forging an invitation. \textit{Position: Controlled. DV 3 (Presence + Performance or Sway).} On a miss, a guard becomes suspicious; advance Social Debt Clock by 1 (someone vouches for you, now you owe them).
    \item \textbf{The Office (Investigation):} Find the vault key or memorize the combination. The office is locked; a riddle on the door: \textit{What grows when given, shrinks when taken, and is never owned?} (Answer: A promise, or a debt). \textit{Position: Desperate if caught. DV 4 (Wits + Lore).} On a success, you bypass the lock. On a miss, you trigger an alarm---but a mysterious masked guest helps you escape; you now owe them a favor (tick Social Debt Clock).
    \item \textbf{The Vault (Climax):} The vault is sealed with a puzzle-lock that requires a \textit{social key}: the correct compliment to a portrait of Lord Caspian's late wife. The answer is carved on the back of a painting elsewhere, but the party can also use a Favor Token with a guest who knows the secret. \textit{Position: Dominant if they have the token; Controlled if they deduce the answer (Wits + Insight DV 4).} Opening the vault incurs no additional debt.
\end{enumerate}

\textbf{Inside the Vault:} The Sunset Ledger is there, along with several magical trinkets. The party also finds a letter that reveals Lord Caspian is himself an Invoker of Livaea---he trades secrets for social standing. The Velvet Countess did not mention this.

\chapter{Act II: The Double-Cross}

\section{Scene 4: The Debt Collector}

The moment the party touches the Ledger, the Social Debt Clock fills (if it hasn't already---GM should time it for drama). The lights flicker. A figure in a porcelain mask steps from behind a curtain: the Debt Collector. It speaks in a soft, androgynous voice: \textit{Who holds the debt for this theft? Name a payer, or the debt becomes a chain.}

The party must decide: \textbf{assign the accumulated debt to one person} (a PC, the Velvet Countess, or an NPC they have a Favor Token with), or \textbf{refuse to pay}---which invokes Livaea's indirect wrath (the Collector disappears, but all future social rolls are at -1 die until the party performs a public humiliation or sacrifice).

\begin{itemize}
    \item \textbf{Assign to NPC:} That NPC will later call in a favor (GM's choice).
    \item \textbf{Assign to Countess:} She nods grimly (\textit{I owe Livaea enough. One more thread won't break me.}) and the party escapes clean.
    \item \textbf{Refuse:} All future social rolls at -1 die until a major sacrifice.
\end{itemize}

\section{Scene 5: The Escape}

The alarm sounds. The party must exit the Villa. They can use remaining Favor Tokens (e.g., a guard they bribed, a window they left open) to reduce the difficulty. \textit{Position: Desperate for any violent approach, Controlled for a planned escape. DV 4 (move, stealth, or command).} On a miss, a party member is captured---but the Velvet Countess arranges their release using her own debt (Social Debt Clock resets but she now owes Livaea a major favor).

\section{Scene 6: The Payoff}

Back at the Velvet Parlor. The Countess examines the Ledger. She burns one page---the evidence against an innocent---and hands the rest to a courier. \textit{Livaea is pleased. Your payment is in the chest.} She also offers each PC a choice: take a magical trinket from the vault (a minor asset), or convert their remaining Favor Tokens into a \textbf{String} (a permanent contact or favor from an NPC). But if they take the trinket, Livaea's interest in them grows. The Velvet Countess warns: \textit{She does not forget. And she does not forgive.}

She then hands them an invitation to the Crimson Court's spring masquerade. \textit{You should attend. It's rude to refuse.}

\chapter{Clocks and Tokens}

\section{Social Debt Clock [6]}

Advances each time a Favor Token is created or spent, or when the party uses a \textit{clever plan} flashback without spending a Boon. When full, the Debt Collector appears.

\begin{infobox}
\textbf{Clock Consequences}
\begin{itemize}
    \item \textbf{Segments 1--3:} Minor inconveniences (awkward invitations, misplaced items).
    \item \textbf{Segments 4--5:} Significant obligations (debts called in during crucial moments).
    \item \textbf{Segment 6:} The Debt Collector appears. A major obligation must be assigned.
\end{itemize}
\end{infobox}

\section{Favor Tokens}

Each is a physical card with an NPC name and a vague obligation. They can be spent to:
\begin{itemize}
    \item Gain +2 dice on a roll related to that NPC's sphere of influence.
    \item Automatically succeed on a minor task (e.g., opening a specific door, learning a password).
    \item Deflect suspicion onto that NPC (costs extra: advance Social Debt Clock by 1).
\end{itemize}

\section{Heist Clocks (Optional)}

\begin{itemize}
    \item \textbf{Alarm [4]:} Advances when the party fails a stealth or social roll. When full, guards swarm; position becomes Desperate for all actions.
    \item \textbf{Suspicion [4]:} Tracks how much Lord Caspian suspects the party specifically. When full, he targets them after the heist.
\end{itemize}

\chapter{Resolution Paths}

\begin{longtable}{p{0.25\textwidth} p{0.7\textwidth}}
\toprule
\textbf{Outcome} & \textbf{Description} \\
\midrule
\textbf{Clean Getaway} & Assign debt to Velvet Countess. Pay 1 Boon each to avoid capture. \\
& \textit{Cost:} Social Debt Clock resets. \\
& \textit{Consequence:} The Countess becomes a recurring contact; Livaea sends a thank-you gift (a cursed handkerchief). \\
\midrule
\textbf{Pass the Debt} & Assign debt to an NPC (e.g., Lord Caspian, a guard). \\
& \textit{Cost:} Lose that Favor Token; GM gains 1 Story Beat. \\
& \textit{Consequence:} That NPC becomes a recurring antagonist or rival. \\
\midrule
\textbf{Burn It All} & Refuse to assign debt. All future social rolls at -1 die until a major sacrifice. \\
& \textit{Cost:} Permanent Condition: \textit{Marked by Livaea} (social penalties). \\
& \textit{Consequence:} Livaea's attention becomes a campaign thread; the Velvet Countess distances herself. \\
\midrule
\textbf{The Heroic Debt} & A PC volunteers to take all accumulated debt personally. \\
& \textit{Cost:} That PC gains a new Persistent Obligation: once per session, GM may force them to accept a social task. \\
& \textit{Consequence:} The PC gains a secret admirer (or stalker) in Livaea's network. \\
\bottomrule
\end{longtable}

\chapter{Handouts and Props}

\section{Favor Tokens}

Small cards with names like \textit{Lord Rencourt -- owes you a night's lodging} or \textit{Mistress Hyssop -- expects a basket of rare herbs.} Players can hold them physically.

\section{The Sunset Ledger}

A prop book with blank pages, but if opened, faint gold ink shows a list of fictional names and bribes.

\section{Invitation to the Crimson Court}

A heavy card with a red wax seal. The text: \textit{Your presence is requested. No need to bring a gift. Your debt is enough.}

\chapter{NPC Profiles}

\begin{longtable}{p{0.25\textwidth} p{0.7\textwidth}}
\toprule
\textbf{The Velvet Countess} & \textbf{Role:} Patron, Invoker \\
& \textbf{Stats:} Presence 4, Wits 3, Spirit 3 \\
& \textbf{Goal:} Balance her own debts while completing the job. \\
& \textbf{Quote:} \textit{I don't serve Livaea out of faith. I serve her because my mortgage is in her name.} \\
\midrule
\textbf{Lord Caspian Verre} & \textbf{Role:} Target, Merchant Prince \\
& \textbf{Stats:} Presence 3, Wits 4, Spirit 2 \\
& \textbf{Goal:} Protect his secrets and social standing. \\
& \textbf{Quote:} \textit{He knows you're here. He just doesn't care---yet.} (Whisper from his portrait) \\
\midrule
\textbf{Mistress Thorn} & \textbf{Role:} Security Chief \\
& \textbf{Stats:} Body 3, Wits 3, Presence 2 \\
& \textbf{Goal:} Test the party's wit before allowing passage. \\
& \textbf{Quote:} \textit{A lock is only as strong as the story behind it.} \\
\midrule
\textbf{The Debt Collector} & \textbf{Role:} Manifestation of Obligation \\
& \textbf{Stats:} Immune to harm. \\
& \textbf{Goal:} Ensure debts are acknowledged. \\
& \textbf{Quote:} \textit{Debt is a story. Who writes the ending---you, or the people you owe?} \\
\bottomrule
\end{longtable}

\chapter{GM Guidance}

\section{Embrace the Parody}

Let players quote heist movie clichés. The humor comes from the juxtaposition of slick plans with mundane social obligations (\textit{I can't pick that lock---I still owe the locksmith a dinner party}).

\section{Livaea is Never Seen}

She is only represented by her agents, her debts, and the velvet curtains that close a little too smoothly. Her power is the power of implication.

\section{The Debt Collector is a Tool}

Use it to apply pressure, not to punish. The Collector might offer deals: \textit{Pay now with a memory, or pay later with a betrayal.}

\section{Favor Tokens Should Be Tangible}

If playing online, use a shared doc. The physical act of giving a token to the GM when spending it increases tension.

\section{The Ending is Open}

The party might decide to attend the Crimson Court masquerade---a perfect lead-in to Livaea-centric political intrigue.

\section{Roleplay Hooks}

\begin{itemize}
    \item \textbf{Velvet Countess:} \textit{I don't serve Livaea out of faith. I serve her because my mortgage is in her name.}
    \item \textbf{Lord Caspian's Portrait:} The eyes seem to follow you. A whisper from nowhere: \textit{He knows you're here. He just doesn't care---yet.}
    \item \textbf{Debt Collector:} \textit{Debt is a story. Who writes the ending---you, or the people you owe?}
    \item \textbf{Party Banter:} \textit{This is just like that time in Zakov!} \textit{No, it's not. That time, we didn't owe a favor to a magpie.}
\end{itemize}

\chapter{Colophon}

The Sunset Ledger now rests in a locked chest beneath the Velvet Parlor. Lord Caspian never reported the theft---he couldn't, without admitting his own debts. The guards who looked away have new watches. The priest who owed a favor now preaches on street corners about the dangers of \textit{ungodly bargains}.

\vspace{1em}
\dusana{And the party? They have a new invitation, a lighter purse, and a handful of debts they don't yet understand. Livaea does not steal. She borrows. And she always, always sends a thank-you note. The ninth candle is for the favors we forget to name.}

\vspace{1em}
\letter{On the road to Silkstrand, with a borrowed horse and a clear conscience.}{--- Dusana of the Raven Road}

\end{document}
