\documentclass[11pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{parskip}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{sectsty}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning}
\usetikzlibrary{shapes,arrows}
\titleformat{\section}{\Large\bfseries\sffamily}{\thesection}{1em}{}
\titleformat{\subsection}{\large\bfseries\sffamily}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\bfseries\sffamily}{\thesubsubsection}{1em}{}
\setlength{\parindent}{0pt}
\setlist{nosep, left=0pt}

\title{The Grumbling Vault \\ A One-Shot for Fate's Edge}
\author{Dusana of the Raven Road \\ (as told by Durin Stoneshield)}
\date{Year of the Iron Harvest}

\begin{document}
\maketitle

\begin{center}
\textit{``You think you know stone? I've drunk ale older than your clan. The Iron Avengers could have held this vault with a teaspoon and a bad attitude. Now? Now you need runes and sparkles. Pathetic.''}
\end{center}

\section*{Introduction}

This one-shot is a pressure cooker for both magic users and melee fighters. Inside an ancient Aeler vault, a volatile keystone pulses with unstable energy. The wards react to every spell, rite, or song, building toward a catastrophic release. Meanwhile, the vault's guardians are built to shrug off magic --- only a well-aimed axe, a hammer blow, or a cracked skull will put them down.

You'll also meet Durin Stoneshield, an Iron Avenger from the old guard. He's bitter, drunk, and full of stories about a past that probably never existed. He's also your only way into the vault.

\textbf{Time:} 3--4 hours \\
\textbf{Players:} 3--5 \\
\textbf{Tier:} II (30--60 XP)

\input{copyright}
\section{Aeler Lore: The Iron Avengers and the Spirit Shield Warriors}

The Aeler don't talk about their failures. They don't pour one out for the dead. They carve the names into stone and move on. But every mountain hold has a locked door, and behind it, old warriors drink alone.

\textbf{Iron Avengers:} Feud-keepers, blood-ledger balancers. They don't forgive; they tally. An Iron Avenger's oath is a debt that outlives generations. Durin was one, once. Now he just complains about the youth.

\textbf{Spirit Shield Warriors:} Crypt-wardens, ancestor-tenders. They fight with ghosts at their shoulders. Their rites are quiet, almost ashamed --- not like the flashy nonsense from southern mages. A Spirit Shield Warrior can turn a blade with the memory of a dead comrade.

Durin claims he was both. No one checks.

\section{The Grumbling Vault}

Deep beneath the Aelerian foothills, a sealed chamber holds the \textbf{Keystone of Unmaking} --- a fist-sized chunk of obsidian that can destabilize any ward within a mile. The Aeler locked it away centuries ago, but the wards have grown senile. They tick and pulse like a failing heart.

Your goal: retrieve the keystone for Durin. He won't say why. He just grumbles about ``old debts.''

\subsection{The Ward Resonance Timer }

The vault reacts to magic. Every time a PC invokes a Rite, sings a Song, performs an Invoker ritual, free casts, or summons a spirit, advance the \textbf{Ward Resonance Timer } (\textbf{8 segments}) by the listed amount. When the timer fills, the \textbf{Ward-Primus} awakens (see below).

\begin{table}[h]
\centering
\begin{tabularx}{\linewidth}{l X}
\toprule
\textbf{Action} & \textbf{Timer Advancement} \\
\midrule
Invoke a Rite (Runekeeper) & +1 \\
Sing a Song (Cantor) & +1 \\
Invoker ritual (normal) & +1 \\
Crack the Seal (Invoker) & +2 \\
Free casting (Spellcraft) & +1 \\
Summon a spirit & +2 \\
Dispel a ward & +1 \\
\midrule
\textit{Suppress a ward (physical action)} & \textit{-1 (can reduce timer)} \\
\textit{Sunder an armored guardian (melee)} & \textit{-1 (see below)} \\
\bottomrule
\end{tabularx}
\end{table}

Whenever the timer advances, the GM gains 1 SB (on top of any from dice) to spend on vault complications: a rune flares (blindness), a stone grinds (difficult terrain), or a minor spirit manifests (Cap 1, attacks the last caster).

\subsection{Melee Challenges: Because Casters Have It Too Easy}

The vault's guardians are \textbf{resistant to magic} (all magical attacks deal -1 Harm, and spells that target ``will'' or ``spirit'' have DV+2). They are, however, \textbf{vulnerable to sundering} --- melee fighters can target their armor plates, chains, or joints.

\begin{itemize}
\item \textbf{Sunder attempt (Body + Melee, DV 4):} On a clean success, reduce the guardian's armor by 1 (to minimum 0) and \textbf{tick the Ward Resonance timer back by 1} (the shattering runes release pent-up energy harmlessly).
\item On a partial, the guardian loses armor but the timer does not retreat. On a miss, the guardian counterattacks (free attack, Harm 1).
\end{itemize}

\subsection{The Old Drunk: Durin Stoneshield}

Durin is an Iron Avenger who retired --- or was pushed out --- decades ago. He wears dented mail that no longer fits, carries a hammer he can't swing, and smells of stale ale and regret. His memory is selective: he remembers glorious battles that never happened, feasts where he was the hero, and oaths that were never sworn.

He accompanies the party into the vault (he knows the way) but is \textbf{useless in combat} (if attacked, he falls prone and curses). His real value: \textbf{Iron Avenger's Insight}. Once per scene, a PC can ask Durin a question about Aeler history, traps, or wards. He will answer with a story that is \textit{mostly} true, but the GM should add one false detail (the party can test with Wits+Lore DV 3 to catch the lie). If they catch it, they gain +1 die on their next action related to that information.

Durin also has a \textbf{Debt Timer (4)}. Each time the party uses his insight or asks him to do something (carry a rope, hold a torch, stop singing), tick it. When it fills, he demands they retrieve a ``trinket'' from the vault --- a third keystone that he claims is his by rights. It's a lie, but refusing will make him storm off, leaving the party trapped.

\textit{(GM: The trinket doesn't exist. If the party searches, they find a rusty belt buckle. Durin accepts it anyway, grumbling about ``close enough.'')}

\section{The Vault: Three Chambers}

\subsection{Chamber 1: The Sealed Door}

A massive bronze door, etched with Aeler runes that pulse faintly. No handle. No keyhole. Durin belches.

\textbf{Traps:} [WARD] --- crossing without dispelling deals 1 Fatigue. [ALARM] --- failure triggers a \textbf{Rune Wisp} (Cap 1, Harm 1).

\textbf{Openings:}
\begin{itemize}
\item \textbf{Crack the door physically (Body + Athletics, DV 5).} On success, door grinds open, no magic used. On failure, advance Ward Timer by 1 (the door resists).
\item \textbf{Dispel the ward (Wits + Arcana, DV 4).} Success opens the door, but advances Timer by 1 (magic was used). Failure advances Timer by 2 and triggers the alarm.
\item \textbf{Ask Durin.} He tells a story about a ``pressure plate under the third stone.'' A Wits+Investigation (DV 3) reveals a loose flagstone. Pulling it opens a crawl space. No magic, no timer advance.
\end{itemize}

\subsection{Chamber 2: The Hall of Sundered Oaths}

A long hall lined with suits of animated Aeler armor --- \textbf{three Iron Wraiths}. They do not move until a spell is cast or a loud noise is made.

\textbf{Iron Wraith} (each):
\begin{itemize}
\item Body 3, Spirit 2, Presence 1
\item Armor 2 (resistant to magic: -1 Harm)
\item Harm 2 (crushing)
\item \textbf{Sunderable:} Melee attacks can reduce armor (see Melee Challenges).
\item \textbf{Death burst:} When destroyed, the wraith explodes in a shower of rust, dealing 1 Fatigue to all in Near range (Body+Endurance DV 3 to avoid).
\end{itemize}

\textbf{Timer advances:} Any magic cast in this room advances Ward Timer by 2 (the runes are sensitive). The party can also intentionally shatter the wraiths' armor to reduce the timer (each sunder attempt that succeeds reduces Timer by 1).

\textbf{Durin's advice (if asked):} ``Back in my day, we'd just kick 'em in the codpiece.'' (This is useless. But if the party laughs, they gain 1 Boon.)

\subsection{Chamber 3: The Heart of Silence}

A domed chamber. The Keystone of Unmaking floats above a black altar. Four pillars pulse with light. The Ward Resonance Timer is visible as a glowing ring around the keystone.

\textbf{The Ward-Primus} (if timer fills) rises from the altar: a construct of living runes, crackling with stored energy.

\begin{itemize}
\item Body 5, Spirit 5, Presence 3
\item Armor 3 (magic immune --- spells deal 0 Harm; only physical weapons work)
\item Harm 3 (force)
\item \textbf{Absorb Magic:} As a reaction to any spell, the Primus negates it and heals 2 Harm.
\item \textbf{Weakness:} Mundane weapons deal normal Harm. Magical weapons suffer -2 dice.
\item \textbf{Death:} When destroyed, the Primus explodes (Harm 2 to all in Near range, Body+Endurance DV 4 to halve).
\end{itemize}

\textbf{If the timer is not yet full}, the party can attempt a \textbf{Stealth Retrieval}:
\begin{itemize}
\item Three successes before two failures using any combination of Stealth, Athletics, Sleight of Hand, or Lore.
\item Each use of magic during this attempt advances Timer by 2 (in addition to any other effects).
\item On three successes, they snatch the keystone and the vault stabilizes. The Primus does not awaken.
\end{itemize}

\textbf{If the timer fills}, the Primus awakens. Fight or negotiate.

\textbf{Negotiation (Presence + Lore, DV 5)}: The Primus can be convinced to release the keystone if the party offers a \textit{different} source of energy --- a memory, a year of their life, or a promise to destroy a rival ward elsewhere. Accepting creates a new Obligation timer (6 segments) for the party.

\section{Endings and Aftermath}

\subsection{Success (Keystone retrieved, Primus dormant or destroyed)}
\begin{itemize}
\item Each PC gains 10 XP.
\item Durin gives them a \textbf{Iron Avenger's Token} (once, reduce Obligation from any Aeler-related Rite by 1) and mutters something about ``not completely useless.''
\item The keystone can be used as a Minor Asset: \textit{Keystone of Unmaking} --- once, suppress all wards in a 30-foot radius for one scene.
\end{itemize}

\subsection{Failure (Keystone lost, Primus rampages, vault collapses)}
\begin{itemize}
\item Each PC gains 6 XP (they survived).
\item They escape with Durin, who blames them for ``not listening to his wisdom.''
\item A new hook: the keystone destabilizes a nearby hold. The party must repair the damage or face the consequences.
\end{itemize}

\subsection{Durin's Debt (if his timer filled)}
He asks for the ``trinket'' --- a rusty belt buckle. If they give it to him, he proclaims them ``honorary Iron Avengers'' and offers a place at his nonexistent table. If they refuse, he storms off and later sends a letter demanding compensation (another hook).

\section*{GM Tips: Running the Grumbling Vault}

\begin{itemize}
\item \textbf{Keep the timer visible.} Use tokens or a whiteboard. Let players see it tick up.
\item \textbf{Make Durin annoying but useful.} His stories are half-true. Reward players who fact-check him.
\item \textbf{Melee matters.} The Iron Wraiths and Primus are designed to frustrate casters. Encourage fighters to sunder and smash.
\item \textbf{SB spending:} Use SB to create environmental hazards: falling rubble (Difficult Terrain), rune flashes (blinded for one round), or a Rune Wisp that harasses the backline.
\item \textbf{Don't be afraid to fill the timer.} The Primus is tough but not unbeatable. A party that relies on magic will face a reckoning.
\end{itemize}

\section*{Colophon}

This adventure was set down by Durin Stoneshield himself, who dictated it between belches. The ink is watered-down ale. The paper is stained. The past he remembers never existed, but the debt is real enough.

Now go. Sunder something.

\end{document}
