\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{array}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{tcolorbox}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{amssymb}

\geometry{a4paper, margin=1in}

\titleformat{\section}{\large\bfseries\scshape}{\thesection}{1em}{}
\titleformat{\subsection}{\bfseries}{\thesubsection}{1em}{}

\pagestyle{fancy}
\fancyhf{}
\lhead{Fate's Edge}
\rhead{The Guest Who Brought Death}
\cfoot{\thepage}

\newtcolorbox{npcbox}[1]{
    colback=gray!5,
    colframe=gray!50,
    fonttitle=\bfseries,
    title=#1,
    boxrule=0.5pt
}

\newtcolorbox{puzzlebox}[1]{
    colback=blue!5,
    colframe=blue!50,
    fonttitle=\bfseries,
    title=#1,
    boxrule=0.5pt
}

\newtcolorbox{cluebox}[1]{
    colback=green!5,
    colframe=green!50,
    fonttitle=\bfseries,
    title=#1,
    boxrule=0.5pt
}

\begin{document}

\begin{center}
    \textbf{\LARGE The Guest Who Brought Death} \\
    \vspace{0.2cm}
    \textbf{Ubral: The Stone Between Spears} \\
    \vspace{0.1cm}
    \small A Fate's Edge Adventure for Tiers I-II (30-70 XP)
\end{center}

\section*{Adventure Overview}
\begin{tabular}{@{}ll}
    \textbf{Region:} & Ubral (The Stone Between Spears) \\
    \textbf{Tiers:} & 1-2 (30-70 XP) \\
    \textbf{Themes:} & Honor, Law, The Price of Oaths, The Weight of Guest-Right \\
    \textbf{Adventure Type:} & Investigation, Social Intrigue, Mystery with Magical Consequences \\
    \textbf{Estimated Play Time:} & 4-6 hours \\
    \textbf{Key Mechanics:} & Positioning, Story Beats, Social Conflict, Resource Management \\
\end{tabular}

\section*{The Hook: Blood on the Hearthstone}
The feast of the First Snow is in full swing at Hearth-Hall, a hill-fort nestled in the high passes of Ubral. The hall is thick with smoke from the central fire, the air warm with the smell of roasting meat and the sound of clashing horns. Ten clans have gathered for the annual Oath-Wedding between House Dorn and House Lorn, where their firstborns will swear a pact of blood and kinship.

The feasting has lasted three days when the host, High Chieftain Varrak the Stone-Tongued, calls for quiet. "By the hearth-stone and the guest-right, I welcome this man to my hall!" He gestures to a weathered traveler with a hooded cloak, who steps forward with a heavy wooden chest.

"Who brings you?" Varrak asks.

"None," the traveler says, his voice rough. "I am my own kin."

Varrak frowns but nods. "Then by the law of the stone, you are welcome here. Share bread and salt."

The traveler is given a seat at the high table. He does not speak much, but his eyes dart about the hall. On the third night, as the feast reaches its height, the traveler stands and raises his tankard. "To the clans of Ubral!" he cries.

The hall erupts in cheers. But as the tankard touches his lips, the traveler gags, clutches his throat, and falls dead. The tankard clatters to the floor, spilling a deep crimson liquid.

The hall falls silent.

A woman screams: "He was poisoned!"

Varrak rises, face grim. "By the hearth-stone and the guest-right, the killer is as safe in this hall as the dead man. None may lay hands upon them, nor speak ill of them, until the cause is known."

The feast has become a prison. The killer is among them, and the clans are two days' ride from their homelands. If the law is broken, blood will flow.

\section*{Act I: The Weight of Law}

\subsection*{Scene 1: The Silent Feast (Social Encounter)}
The PCs arrive as travelers seeking shelter for the snow, envoys sent to witness the Oath-Wedding, or hunters caught in the blizzard. They are welcomed into the hall and given the grim news. If they ask to leave, Varrak warns: "You saw the death. The law binds you as it binds us."

\begin{itemize}
    \item \textbf{Objectives:} Gain trust with the clans, discover who had reason to kill the traveler, learn why the traveler was here
    \item \textbf{Position:} Controlled (they are guests)
    \item \textbf{DV:} 2 (tense atmosphere; DV 3 for high-stakes information)
    \item \textbf{Consequences:} Failed rolls against DV 3 grant 1 Story Beat to the opposing NPC. Failed rolls against DV 2 make the NPC wary (next roll +1 DV).
\end{itemize}

\begin{npcbox}{Key NPCs}
    \textbf{Brin of House Dorn (young, anxious):} "My cousin was to wed. Now my house may burn." \\
    \textbf{Kael of House Lorn (bitter):} "The traveler was to ruin us. He was not welcome." \\
    \textbf{Bran the Stone-Reader (village elder):} "The hearth-stone sings of blood, but not from a traitor."
\end{npcbox}

\begin{puzzlebox}{The Traveler's Chest}
The traveler's chest is locked with a simple mechanism that requires knowledge of Ubral's oaths.
\begin{itemize}
    \item \textbf{Mechanic:} Wits + Lore vs DV 3 (to solve the lock)
    \item \textbf{Inside the chest:} A map of the high passes with a circled location, a ledger of names, and a silver key
\end{itemize}
\end{puzzlebox}

\begin{cluebox}{Clue}
The ledger lists "oaths to be broken" - names of clan elders who have broken sacred oaths. The map shows the location of the Soul-Snare.
\end{cluebox}

\subsection*{Scene 2: The Broken Oath (Puzzle Encounter)}
The PCs discover the traveler was an Oath-Breaker - a member of a secret society that punishes oath-breakers.

\begin{itemize}
    \item The traveler's body is found with a broken stone ring, the symbol of the Oath-Breakers
    \item The killer took one fragment of the ring
\end{itemize}

\begin{puzzlebox}{Reassemble the Oath-Breaker's Ring}
The ring must be reassembled to learn the traveler's purpose.
\begin{itemize}
    \item \textbf{Puzzle Mechanics:} Five fragments must be found throughout the hall and placed in the correct order based on the Oath of Stone and Spear:
    \begin{enumerate}
        \item Stone (Stability) - Fragment found near hearth-stone
        \item Spear (Defense) - Fragment found near weapons rack
        \item Hearth (Kin) - Fragment found near wedding banners
        \item Sky (Oath) - Fragment found near carved sky-gods
        \item Earth (Return) - Fragment found near burial cairn window
    \end{enumerate}
    \item \textbf{Solving reveals:} The traveler was an Oath-Breaker sent to enforce oaths
    \item \textbf{Failure:} Failing to solve within 2 hours results in 1 Story Beat and a false lead (e.g., House Lorn is the primary oath-breaker)
\end{itemize}
\end{puzzlebox}

\section*{Act II: The Blood-Price Rises}

\subsection*{Scene 3: The First Blood (Combat Encounter)}
The next day, a blizzard traps everyone inside. The traveler's body vanishes, and a ritual is being performed outside.

\begin{itemize}
    \item \textbf{Antagonist:} Lira of House Dorn (a young woman driven mad by visions)
    \item \textbf{Guardian:} Wraith of Broken Oath (a shadowy figure that attacks with spectral blades)
    \item \textbf{Setting:} The burial cairn outside Hearth-Hall, with glowing runes etched in the snow
\end{itemize}

\begin{puzzlebox}{Runes of the Cairn}
\begin{itemize}
    \item The Wraith is immune to physical harm unless the runes are broken
    \item Stepping on a rune inflicts 1 Fatigue and makes you visible to the Wraith
    \item A successful Wits + Awareness roll (DV 2) allows a character to spot the faintly glowing rune before stepping on it
    \item Breaking a rune (Wits + Lore, DV 3) reveals one memory of the traveler
    \item The runes must be broken in the correct order (based on Ubral's oaths)
\end{itemize}
\end{puzzlebox}

\begin{cluebox}{Clue}
One memory reveals the traveler was not an Oath-Breaker, but a Ward-Reader sent to protect the Oath-Wedding.
\end{cluebox}

\subsection*{Scene 4: The Warden of Oaths (Magical Encounter)}
The PCs find a journal that reveals the true purpose of the traveler.

\begin{puzzlebox}{The Journal of Oaths}
\begin{itemize}
    \item \textbf{Puzzle:} The journal is protected by a Ward of Oath-Keeping
    \item \textbf{Mechanics:} One PC at a time may attempt to read (Wits + Spirit vs DV 3)
    \item \textbf{Consequences:} Each failure inflicts 1 Fatigue and reveals a half-truth. A character can attempt again after a brief rest (10 mins, 1 SB spent), reducing the Fatigue cost by 1 for that attempt.
\end{itemize}
\end{puzzlebox}

\begin{cluebox}{Revealed Truth}
\begin{itemize}
    \item The traveler was sent by a Ward-Reader to protect the Oath-Wedding from a Soul-Snare
    \item The Soul-Snare is a magical trap that would have bound the clans to eternal enmity
    \item The true traitor is still among them, and the Soul-Snare is active
\end{itemize}
\end{cluebox}

\section*{Act III: The Oath That Binds}

\subsection*{Scene 5: The True Traitor (Social Climax)}
The true traitor is Bran the Stone-Reader, who has been feeding false information to Lira. He believed the Oath-Wedding was a trap by House Lorn. The Soul-Snare is active and will spring at dawn.

\textbf{Final Choice:}
\begin{itemize}
    \item \textbf{Confront Bran directly:} Requires a successful social contest (Wits + Presence vs DV 3)
    \item \textbf{Let him expose himself:} Requires the PCs to steer events (e.g., convincing Lira to recount visions under oath, planting evidence, publically reading journal half-truths)
\end{itemize}

\subsection*{Scene 6: The Breaking of Oaths (Climactic Confrontation)}
The true threat arrives - the Ward-Reader, a figure cloaked in shifting stone.

\begin{itemize}
    \item \textbf{Antagonist:} The Ward-Reader (uses earth and stone magic)
    \item \textbf{Guardians:} Two Oath-Bound Guardians (spectral figures that attack with the force of broken promises)
\end{itemize}

\textbf{Climactic Choice:}
\begin{itemize}
    \item \textbf{Let the Soul-Snare consume someone:} The Oath-Wedding is completed at great cost
    \item \textbf{Try to break the Soul-Snare without sacrifice:} Risk of failure
    \item \textbf{Let the trap activate:} The clans are bound to eternal enmity
\end{itemize}

\section*{Consequences}

\begin{itemize}
    \item \textbf{If the Soul-Snare is Broken Without Sacrifice:}
    \begin{itemize}
        \item The Oath-Wedding is completed, but the cost is high - one PC or NPC is cursed with "The Weight of Oaths" (1 Fatigue penalty for all social rolls for 1 week)
        \item The Ward-Reader is banished, but swears revenge
    \end{itemize}
    
    \item \textbf{If a Sacrifice is Made:}
    \begin{itemize}
        \item The Oath-Wedding is completed
        \item The sacrificed person becomes a legend in Ubral
        \item The Ward-Reader is banished
        \item The PCs gain the respect of the clans (1 level of Social Capital with Ubral)
    \end{itemize}
    
    \item \textbf{If the Soul-Snare is Not Broken:}
    \begin{itemize}
        \item The clans go to war
        \item The PCs are branded oath-breakers
        \item They are exiled from Ubral
    \end{itemize}
\end{itemize}

\section*{GM Guidance}

\textbf{Ubral Cultural Elements to Emphasize:}
\begin{itemize}
    \item Guest-right is sacred: even the killer is protected until the truth is known
    \item Every action has weight: oaths are binding, and breaking them has consequences
    \item The land remembers: broken oaths are remembered by the stones
\end{itemize}

\textbf{Pacing Tips:}
\begin{itemize}
    \item Let the social encounters build tension before the magical combat
    \item Verbally track time pressure: "45 minutes gone... the wind is picking up outside..."
    \item Make the Soul-Snare a real threat - have the PCs feel the clock ticking
    \item The final choice should feel momentous, with no "perfect" answer
\end{itemize}

\textbf{Themes to Highlight:}
\begin{itemize}
    \item The cost of law and order
    \item How traditions can both bind and protect
    \item The weight of choice in a world where oaths shape reality
\end{itemize}

\section*{Adventure Seeds for Future Sessions}
\begin{itemize}
    \item \textbf{The Weight of Oaths:} The PC who made the sacrifice returns as a Wraith of Broken Oath.
    \item \textbf{The Ward-Reader's Revenge:} The exiled Ward-Reader gathers forces to break the Oath-Wedding.
    \item \textbf{The Unbroken Oath:} The clans demand the PCs mediate a dispute between warring clans.
\end{itemize}

\section*{GM Notes}
This adventure is a story about the weight of law and the cost of oaths. In Ubral, every word matters, every action echoes, and the land itself remembers. The adventure forces the PCs to confront the true price of the laws they live by—and decide what they are willing to break to save a people who believe in laws that may be their ruin.

May your choices be as heavy as the stones of Ubral.

\end{document}