\documentclass[10pt,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{paper=b5, margin=1.2in, top=1.5in, bottom=1.5in}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{titletoc}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage[hang]{footmisc}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    linkcolor=black,
    citecolor=black,
    urlcolor=black,
}

\definecolor{chainGrey}{RGB}{80,80,80}
\definecolor{threadRed}{RGB}{160,40,40}
\definecolor{ichorBlack}{RGB}{20,20,25}
\definecolor{wornGreen}{RGB}{70,90,60}
\definecolor{parchment}{RGB}{245,240,225}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[RE,LO]{\textit{The Hollowed Hunger / The Worn Tide}}
\fancyhead[LE,RO]{\thepage}
\fancyfoot[CE,CO]{\textcolor{chainGrey}{--- \textit{The debt is never closed. It only changes shape.} ---}}

\titleformat{\chapter}[display]
  {\normalfont\Huge\bfseries\color{threadRed}}
  {\titlerule[1pt]\vspace{4pt}\Large\chaptertitlename\ \thechapter}
  {0pt}
  {}
\titlespacing*{\chapter}{0pt}{-30pt}{20pt}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\newenvironment{statblock}
  {\begin{center}\begin{minipage}{0.85\textwidth}\small\setlength{\parskip}{0.3em}\begin{quote}}
  {\end{quote}\end{minipage}\end{center}}

\begin{document}

\titlepage
\begin{center}
\vspace*{2cm}
{\huge\bfseries\color{threadRed} The Hollowed Hunger}\\[0.8cm]
{\Large\itshape A Sequel to the Hollowed Plague}\\[1.5cm]
\rule{0.6\textwidth}{1pt}\\[0.5cm]
\emph{An Epic Adventure for Tier IV--V Characters}\\[1cm]
\vfill
{\small \textcolor{chainGrey}{Under the lamp, bread and salt --- and a hunger that cannot be named.}}\\[0.2cm]
{\small \textcolor{chainGrey}{The ink is fresh. The road forks again. The debt is never closed.}}
\end{center}
\newpage

\chapter*{Safety and Consent at the Table}
\addcontentsline{toc}{chapter}{Safety and Consent at the Table}
This adventure includes themes of mass suffering, famine, undeath, body horror, and difficult moral choices. It directly follows the events of \textit{The Hollowed Plague} and assumes that the party has already witnessed the horror of the Turned Hollowed.

Before play, establish:
\begin{itemize}
\item \textbf{Lines:} Hard boundaries. Example: no sexual violence, no harm to children (the Hollowed are not human children), no graphic torture.
\item \textbf{Veils:} Soft boundaries. Example: the hunger of the Hollowed can be described obliquely rather than depicted in detail.
\item \textbf{X‑Card:} Any player may tap an X card to pause or remove a scene without explanation.
\end{itemize}
Use these tools.

\tableofcontents
\newpage

\chapter{Prologue -- What Came Before}

\section{Determining the Branch}
\textit{The Hollowed Plague} ended with one of three outcomes. This sequel adapts to each.

Ask your table (or recall):
\begin{itemize}
\item \textbf{Which path did the party take?}
  \begin{itemize}
  \item \textbf{The First Needle (Chain Path):} The party retrieved the First Needle and ended the plague, returning the Turned Hollowed to their silent, heartless state. Malachai was thwarted -- but she used the party's intrusion into Zahk"Kozahn as a catalyst to tear open a rift. The \textbf{Worn Tide} begins.
  \item \textbf{The Flash of Ichor (Malachai Path):} The party stole the Ichor and gave the Turned Hollowed true sentience. They are free, but a new hunger grows in them -- a thirst for truth, for restitution, for the debts they witnessed. The \textbf{Hollowed Hunger} begins.
  \item \textbf{Doing Nothing (Catastrophe):} The party delayed; the Cord broke. The Hollowed scattered across every threshold, becoming a permanent plague. Nations burn. This branch is the darkest timeline, and is covered in Appendix C.
  \end{itemize}
\end{itemize}

\section{The World After}
Regardless of the branch, the region is scarred. The Daughters of the Unbroken Cord are fractured. Malachai is more active than ever. And the party is known -- as saviours, as liberators, or as those who failed to choose.

\chapter{Common Act I -- The Scarred Land}

\section{The Hook}
Months have passed. The party receives a summons -- or a threat -- depending on the branch.

\begin{itemize}
\item \textbf{Worn Tide:} A massive wave of ghouls (\textit{Worn}) has spilled from the desert, overwhelming frontier garrisons. A desperate messenger from the Aeler holds begs for aid. The Worn are not mindless; they are organised, led by figures who wear the remnants of Daughters" robes.
\item \textbf{Hollowed Hunger:} Villages along the old caravan routes report strange visitors -- children with silver wire eyes who do not age, do not eat, and ask only one question: "What do you owe?" Some offer clemency. Others extract payment in blood.
\end{itemize}

The party is the only group with knowledge of both factions. They must intervene.

\section{Key NPCs -- Act I}
\begin{itemize}
\item \textbf{Sister Vorna (Winter of the Chain, Wounded):} She survived the plague but lost her coven. She now wanders, offering what aid she can. She is bitter, but she trusts the party. \textit{"I thought the Needle was mercy. Now I wonder if mercy is just another word for control."}
\item \textbf{The First Free (Hollowed, named Ellia):} The first Hollowed to receive the Ichor. She is now their reluctant leader. She does not want to hunger. She does not want to judge. But she has no choice. \textit{"I see every debt. Every unkind word. Every broken promise. I did not ask for this sight."}
\item \textbf{Collector Prime (The Silent One who turned):} A former Hollowed who embraced Malachai's curse. It now serves as a liaison between the Worn and the Chained Angel. It speaks in flat, echoing tones. \textit{"The Well is dry. The thread is cut. Only hunger remains."}
\end{itemize}

\section{Key Encounters -- Act I}

\subsection{The Rift or the Hunger (Investigation)}
The party must learn the scope of the new threat. Use the travel deck (Acasia, Dhahara, or Aeler) to generate scene seeds. Common clues:
\begin{itemize}
\item A village where every inhabitant has been judged by a Hollowed -- some spared, some executed.
\item A fortress overrun by Worn, with a strange banner bearing Malachai's ship sigil.
\item A Daughter's hospice converted into a shrine to the Chained Angel.
\end{itemize}

A successful Wits + Lore (DV 4) reveals that the source of both threats lies in the same place: \textbf{the threshold where the Cord was broken (House Without Doors) or the rift where the party exited Zahk"Kozahn}. The party must return.

\subsection{The Faction's Plea (Social)}
Sister Vorna offers the party a fragile alliance with the surviving Daughters. In exchange for help, she can provide healing, shelter, and a few remaining loyal Hollowed (not turned). But she warns: "The High Mother is silent. The Cord is unstable. We cannot help you fight -- only survive."

If the party seeks aid from the Hollowed (branch dependent), they must negotiate with Ellia or the Collector Prime. The price is always a debt: a promise to witness a future oath, to spare a certain enemy, or to surrender a memory.

\subsection{The First Incursion (Combat or Defense)}
A band of Worn (if Worn Tide) or a pack of Hungry Hollowed (if Hollowed Hunger) attacks a settlement the party is in. This is a tutorial combat to establish the new threat.

\textbf{Worn (Ghouls):} Cap 2, Tier III. Abilities: \textit{Ichor Hunger} (feed on corpse to heal), \textit{Desperate Lunge} (extra action when bloodied). Weakness: Fire. Led by a \textbf{Worn Captain} (Cap 4, Tier IV) who wears a Daughter's grey robe as a trophy.

\textbf{Hungry Hollowed:} Cap 3, Tier III. Abilities: \textit{Silver Gaze} (force truth), \textit{Witness Recall} (reveal a secret about the target). Weakness: Cannot cross a threshold of running water.

\section{Act I Resolution}
The party learns that the source of the crisis is a broken threshold -- the place where the Cord snapped (if Hollowed Hunger) or the rift Malachai tore (if Worn Tide). They must travel there and seal it. Two paths emerge, mirroring the fork of the previous adventure but inverted.

\chapter{Act II -- The Fork Inverted}

\section{Path A -- The Worn Tide (Repairing the Rift)}

\subsection{Return to Zahk"Kozahn (Requirement)}
The rift is located in the cursed city. The party must return -- but this time, Malachai's defences are heightened. The city is now openly hostile.

\subsubsection{Step 1 -- The Worn Siege Line}
The approach to Zahk"Kozahn is blocked by a massive encampment of Worn, organised into warbands. The party must infiltrate or fight through.

\textbf{Infiltrate:} Wits + Stealth, DV 5, using captured Worn robes (requires a previous kill). Failure triggers a massive chase.

\textbf{Fight:} A multi‑stage battle. Timers: \textbf{Worn Morale [8]} -- ticks down as the party kills leaders. At 0, the warbands break.

\subsubsection{Step 2 -- The Cathedral Under Siege}
The rift is in the cathedral where Malachai is chained. She is now partially free -- one of her arms has broken the Aetherium chain. She cannot leave, but she can direct the Worn.

The party must fight through the cathedral's defenders (Worn and a few Turned Hollowed who serve her) to reach the rift.

\subsubsection{Step 3 -- Sealing the Rift (Final Encounter)}
The rift is a tear in reality, bleeding ichor and Worn. To seal it, the party needs a \textbf{Thread of the Cord} -- a fragment of the High Mother's power. Sister Vorna provides one, but it must be woven into the rift.

\textbf{Ritual Sealing:} Requires three successes (Arcana, Lore, or Presence) before two failures. On each failure, a Wave of Worn attacks (Cap 3).

If successful, the rift closes. Malachai screams. Her arm remains free, but the tide is stemmed. The party escapes as the city begins to collapse.

\subsection{Outcome -- Worn Tide Resolved}
The Worn are cut off from reinforcements. The remnants can be hunted down. Malachai is weakened. The Daughters begin to rebuild, but the High Mother is near death. The party gains a permanent Asset: \textbf{Rift Shard} (once per session, banish a single undead creature).

\section{Path B -- The Hollowed Hunger (Reweaving the Covenant)}

\subsection{Return to the House Without Doors}
The Hungry Hollowed are drawn to the place where they were first unmade -- the House Without Doors. The party must enter again, but the House has changed. It is now guarded by Hollowed who have not turned -- and by those who have.

\subsubsection{Step 1 -- The Liminal Descent}
The threshold trials are inverted. Instead of proving loyalty to the Chain, the party must prove that they respect the Hollowed's new freedom.

\textbf{Trial of the Witness:} A Hollowed asks each PC: "What debt have you paid that you did not owe?" An honest answer (roleplayed) grants passage. A lie or evasion summons a Hungry Hollowed.

\textbf{Trial of the Unbroken Cord:} The cord is now frayed. The party must decide whether to cut it (freeing the House entirely) or HEAL it (restoring the Chain's control). Cutting gives the Hollowed true autonomy but removes the Chain's protection. Mending restores the old order -- but enslaves the Hollowed again.

\subsubsection{Step 2 -- The High Mother's Judgment}
The High Mother is in agony, pulled between the needs of the Cord and the demands of the freed Hollowed. She offers a choice: "Help me reweave the Cord, and the Hollowed will return to silence. Or help them cut it, and I will die. The Hollowed will be free -- but they will hunger forever."

The party must choose. There is no compromise.

\subsubsection{Step 3 -- The Final Choice (No Roll)}
If the party mends the Cord, the Hollowed lose their sentience. They return to being silent witnesses. The High Mother survives, weakened. The party gains a String: \textbf{Chain's Gratitude} (once per session, call on a Daughter for aid). But the Hollowed who tasted freedom will remember. Some will turn again.

If the party cuts the Cord, the High Mother dies. The House Without Doors collapses. The Hollowed are free -- but they are now a new people, bound by no law, hungry for truth. They may become allies or enemies. The party gains a String: \textbf{Hollowed Covenant} (once per session, ask a Hollowed for a witnessed secret). But the Cord is destroyed, and the Daughters scatter.

\subsection{Outcome -- Hollowed Hunger Resolved}
The hunger does not end. It changes. The Hollowed become a permanent fixture of the world -- sometimes merciful, sometimes terrible. The party's choice defines their relationship with them forever.

\chapter{Act III -- The Debt That Remains}

\section{Epilogue Scenes}
Depending on the branch and the choices made, the GM narrates a montage of consequences.

\subsection{Worn Tide -- Aftermath}
\begin{itemize}
\item The Worn are pushed back to the desert. Malachai remains chained, but one arm is free. She can now send her influence further.
\item The Aeler holds honour the party as saviours. They offer a permanent hold in the mountains.
\item Sister Vorna becomes the new High Mother -- a reluctant, grieving leader.
\end{itemize}

\subsection{Hollowed Hunger -- Aftermath}
\begin{itemize}
\item The Hollowed establish their own settlements -- silent villages where the inhabitants never age and never forget.
\item Some humans seek them out, hoping to have their debts judged and cleared.
\item Others hunt them, fearing what they represent.
\item The party is caught between, known as the ones who gave freedom -- or the ones who tried to take it away.
\end{itemize}

\subsection{Catastrophe Branch -- The Scattered Witnesses}
If the party did nothing in the previous adventure (see Appendix C), the Hollowed are everywhere. The party must now gather the fragments of the Cord and make a choice -- end the plague by re‑enslaving them, or let them continue as a haunting presence. This branch is shorter and darker, intended as a tragedy.

\chapter{Appendix A -- GM Tools}

\section{Campaign Timers}
\begin{itemize}
\item \textbf{Worn Tide [10]} -- How widespread the ghoul invasion has become. Fills as the party delays. At 8, major cities fall.
\item \textbf{Hollowed Hunger [10]} -- How desperate the freed Hollowed have become. Fills as they are hunted or denied restitution. At 8, they begin to judge indiscriminately.
\end{itemize}

\section{Key NPC Stat Blocks}
\subsection{Worn Captain (Cap 4, Tier IV)}
\textit{Abilities:} Ichor Burst (area damage), Command Worn (summon 2 Cap 2 ghouls), Unnatural Vitality (regenerate 1 Harm per round). \textit{Weakness:} Decapitation.

\subsection{Hungry Hollowed Judge (Cap 3, Tier III--IV)}
\textit{Abilities:} Silver Gaze (compel truth), Debt Collection (target loses 1 Boon or takes 2 Harm), Witness Summon (call a nearby Hollowed). \textit{Weakness:} Cannot harm someone who has publicly confessed a debt.

\section{Random Complication Table (d12)}
\begin{enumerate}
\item A Worn captain recognises a PC from the previous adventure.
\item A Hungry Hollowed demands a PC's true name as payment for passage.
\item The rift pulses; a wave of ichor floods the chamber.
\item A Daughter, now turned, offers the party a deal -- help her escape in exchange for information.
\item The High Mother's whispers guide the party -- but also demand a memory.
\item Malachai's chained hand reaches out; a PC must test Resolve DV 5 or be cursed.
\item A Hollowed child appears and asks: "Why did you wake us?"
\item The floor collapses into a nest of Worn eggs (ticking timer).
\item A rival faction (Ecktoria, Vilikari, Aeler) arrives to claim the rift for themselves.
\item The party's own reflections in ichor show them as Hollowed.
\item A Silent One (not turned) offers to become a permanent witness for the party -- at a price.
\item The Cord reforms, but it is black. It now serves something else.
\end{enumerate}

\chapter{Appendix B -- Pre‑Generated Character Updates}
For parties continuing from \textit{The Hollowed Plague}, award 15--20 XP and suggest:

\begin{itemize}
\item \textbf{New Talent: Ichor Resilience} -- ignore first poison or disease effect per scene.
\item \textbf{New Talent: Witness of the Cord} -- once per session, ask the GM one question about a debt or oath.
\item \textbf{New Asset: Hollowed Contact} -- a loyal Hollowed (Cap 2) who provides information.
\end{itemize}

\chapter{Appendix C -- The Catastrophe Branch (Doing Nothing)}
If the party did nothing in \textit{The Hollowed Plague}, the Cord broke. The Hollowed are scattered, and every threshold is a potential witness.

\section{The Hook}
The party is summoned by a desperate coalition of nobles, merchants, and clergy. The Hollowed appear everywhere -- at births, deaths, weddings, and treaties. They do not attack; they simply watch. And everyone knows that they are judging.

\section{The Resolution}
The party must gather the fragments of the Cord (scattered across the region) and either:
\begin{itemize}
\item \textbf{Re‑weave the Cord:} Re‑enslave the Hollowed, ending the haunting but committing a terrible crime.
\item \textbf{Destroy the Cord forever:} The Hollowed remain free, but without any anchor. They become ghosts -- silent, harmless, but everywhere.
\item \textbf{Broker a Covenant:} Negotiate a new agreement between the living and the Hollowed, giving the Hollowed a place in society. Very high DV, requires extraordinary roleplay.
\end{itemize}
This branch is intended as a bittersweet, tragic ending.

\chapter{Colophon}
\addcontentsline{toc}{chapter}{Colophon}
This adventure was written on the road between hospices, between wars, between hungers. I spoke to a Daughter who had lost her Cord, and to a Hollowed who remembered her own death.

The thread is frayed. The rift is open. The debt is never closed.

\vspace{0.5cm}
\noindent
\textit{The ink is fresh. The road forks again. Choose.}

\end{document}
