\documentclass[11pt]{article}

%=========
% Encoding & Layout
%=========
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{parskip}

%=========
% Fonts & Formatting
%=========
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}

%=========
% Color & Boxes
%=========
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}

\definecolor{mistgrey}{RGB}{180,190,200}
\definecolor{diregreen}{RGB}{23,66,48}
\definecolor{bloodrust}{RGB}{112,32,32}

\tcbset{
  colback=mixwhite!98!mistgrey,
  colframe=diregreen,
  sharp corners,
  boxrule=0.7pt
}

%=========
% Hyperlinks
%=========
\usepackage[hidelinks]{hyperref}

% Optional: quick tag macro if you"re using [TAGS]
≠wcommand{\Tag}[1]{\textsf{\textbf{[}#1\textbf{]}}}

%=========
% Title
%=========
\title{\vspace{−1.5cm}\bfseries\Huge Into the Direwood\\[4pt]
Łarge A Tier~III−−IV Survival Horror Saga}
\author{łarge A Fate's Edge Adventure Module}
\date{\small Version 0.9 −− →day}

\begin{document}
\maketitle
þispagestyle{empty}

% Optional mood blurb on title page
\begin{tcolorbox}
¢ering
\emph{Beyond the last bell−line, where the mist never lifts and the roots drink only memory,\\
an ancient spear pins a dead druid to a fortress of bone.\\
Pull it free, and the forest remembers how to scream.}
\end{tcolorbox}

\clearpage
\tableofcontents
\clearpage

%=== Adventure text begins here ===

\section{Into the Direwood (Tier III−−IV Adventure)}
łabel{sec:into−direwood}

\subsection*{Premise and Tone}

\begin{itemize}
  ℹtem \textbf{Recommended Tier:} III−−IV (experienced crews)
  ℹtem \textbf{Atmosphere:} Half cosmic horror, half survival horror
  ℹtem \textbf{Themes:} Lost names, corrupted guardianship, identity erased by history, truth vs.\ myth
\end{itemize}

The Direwood is a blighted frontier forest where an ancient fortress vomits mist and undead into the world. Long ago, a nameless champion impaled a corrupted druid with a \textbf{gilden spear}−−−but the act unleashed the Mist that devoured her, erased her name, and warped her soul into a tragic antagonist.

Deep within the Direwood lies the fortress heart where she and the druid remain bound in a cycle of torment. Heroes must enter, survive, and uncover the truth:  
\textbf{The curse persists not because the druid lives, but because the heroine's name was erased.}

Only by revealing her name and restoring her saga can the curse be broken.

%====================================================
\subsection{Setup and Hooks}

This adventure is \textbf{setting−agnostic}. Any region bordering cursed wilderness can use it.

⫕section*{Universal Hooks}

\begin{itemize}
  ℹtem \textbf{The Mist Thickens:} Something wakes in the Direwood. Fog breaches warding lines−−−new horrors stalk settlements.
  ℹtem \textbf{A Fragment of Saga:} Scholars, oracles, or ravings speak of a "nameless champion" and a spear of gilded dawn.
  ℹtem \textbf{Faith or Faction Interest:} Reformers, pilgrims, war cults, or occult orders want the fortress breached or sanctified.
  ℹtem \textbf{A Bargain:} A ghostly figure visits a character in a dream:  
        \emph{"Find my name, free me... or drown with me."}
\end{itemize}

You may replace factions with whatever religions, orders, or political blocs exist in your campaign.

%====================================================
\subsection{Historical Truth}

Once, a powerful druid served as the Direwood's warden.  
When an exiled necromancer befriended him, their work twisted root and bone into death−magic. The druid raised undead armies−−−Shades, Mistwraiths, Shadewalkers, and beast−things.

A champion−−−her culture forgotten−−−defeated him using a \textbf{gilden spear} of unknown origin.

But victory was a lie.

\begin{itemize}
  ℹtem The druid's death ruptured a dimensional wound.
  ℹtem Mist poured out, consuming the heroine and her companions.
  ℹtem She became a bound shade−−−hero and warden both−−−now twisted into a \textbf{Guardian of Silence}.
\end{itemize}

The world remembered the druid as the villain but \emph{forgot her name}.  
That erasure is the core of the curse.

%====================================================
\subsection{Adventure Structure}

Run this as a 3−−5 act ordeal:

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
%====================================================
⫕section*{Act I: Payden's Port and the Forgotten Chord}

Tone: political gothic + creeping horror

\paragraph{Premise:}
The warding bell−line falters. Bells toll out of sequence,  
Mist incursions grow bolder, and refugees whisper of shapes in fog.

The PCs arrive in Payden's Port −−− a sea−garrison city of cracked statues,  
storm−beaten docks, and chanting wardens −−− and are pulled into crisis.

\paragraph{The Hook:}
Only one thing can repair the failing bell−line −−−  
\textbf{the Chord of the First Curse}:  
a strip of woven gold and root, a relic that was  
\emph{torn away and lost in the Direwood when the first curse was sealed}.

Legends say it once bridged sound and spirit, harmonizing  
the ward with the heroine's true name.

\paragraph{Act Beats}

\begin{itemize}
  ℹtem \textbf{Factional pressure}:  
  The Aeler garrison, Mistlander priests, and Guild factors all demand different outcomes; each tries to recruit or manipulate the PCs.
  
  ℹtem \textbf{Living omens}:  
  Dead gulls arrange themselves in votive circles, fog−children whisper names not spoken.
  
  ℹtem \textbf{The first breach:}  
  The city's outer bell−line sputters; Mistwraiths claw at the docks. PCs must assist the wardens or improvise defenses −−− gaining first taste of the Mist.
  
  ℹtem \textbf{Call to action:}  
  Divination, confession, or coercion leads to one truth:
  \emph{the Chord lies deep in the Direwood,  
  near the fortress where the heroine fell.}
  
  ℹtem \textbf{The Charge}:  
  "Retrieve the Chord and restore the bell−line −−−  
  or the Port will drown in white."
  
  The PCs are sent through the breached ward with blessings, suspicions,  
  or threats at their back.
\end{itemize}

\paragraph{Dread Timer }
% moved unchanged, still applicable
Mark ticks when:
\begin{itemize}
    ℹtem the party ignores port omens or visions
    ℹtem they mishandle the dead or desecrate wards
    ℹtem they underestimate or bargain with Mist
\end{itemize}

If filled early, Act II becomes harsher and  
the heroine initially appears as a hostile nemesis.

\vspace{0.5em}

\noindent\textbf{Act I Exit Conditions:}
\begin{itemize}
    ℹtem the party crosses into the Direwood,
    ℹtem an alliance (or enmity) with at least one faction is set,
    ℹtem the first imprint of the nameless heroine haunts one PC.
\end{itemize}

⫕section*{Act II: Through the Direwood}

\textbf{Tone:} Survival horror, mythic fever−dream, predatory landscape.

Crossing the tree−line is not entering a forest−−−it is passing into a place shaped by memory, undeath, and betrayal. The Direwood resists mapping; it \emph{reacts} to intrusion.

\paragraph{Environmental Themes}
\begin{itemize}
  ℹtem \textbf{Twisting Trails:} Paths repeat; distances elongate. A day's march may loop back to a familiar tree, now adorned with new corpses.
  ℹtem \textbf{Living Landmarks:} Hollow oaks bleed pale sap, stone cairns whisper warnings, ravens speak in half−remembered curses.
  ℹtem \textbf{Memory Echoes:} Scenes replay−−−villagers fleeing, army banners burning, lovers separated−−−yet the faces blur and distort.
  ℹtem \textbf{Dream Fatigue:} Rest is possible, but sleep is disturbed. PCs wake with soil under their nails, leaves in their hair, or frost upon their cheeks.
\end{itemize}

\paragraph{Adversaries and Threats}
\begin{description}
  ℹtem[Mistwraith Packs:] \emph{Tier III} predators that sense rising fatigue; they do not eat−−−\emph{they erase}.  
    PCs marked by fear or exhaustion attract them.
  ℹtem[Shadewalkers:] \emph{Tier II−III} revenant infantry−−−stone−cold tacticians reenacting endless patrols until interrupted.
  ℹtem[Hunger Wolves:] \emph{Tier III beast−spirits}, sinew stretched over bone, jaws unhinging to swallow light itself.
  ℹtem[The Forgotten Chorus:] \emph{Tier IV psychic hazard}−−−mourning voices that harmonize grief into psychic drowning.
  ℹtem[The Heroine's Phantom:] A \emph{shifting Tier IV entity}−−−sometimes guide, sometimes executioner. Her mood tracks the party's reverence for lost history.
\end{description}

\paragraph{Nightly Visitations}
Each night, one PC dreams of the nameless heroine:
\begin{itemize}
    ℹtem Meeting her mother beneath a moonlit birch.
    ℹtem Riding with oath−sworn shield−sisters.
    ℹtem Walking toward fire knowing she will die.
\end{itemize}

\textbf{GM Note:} Fragments arrive out of order, wrong names inserted−−−only when her true name is spoken do the visions align.

\paragraph{Direwood Survival Procedures}
\begin{enumerate}
  ℹtem At each travel segment, roll the \textbf{Direwood Consequence Table}.
  ℹtem Reduce safe camp options−−−firelight attracts attention, cold invites wraiths.
  ℹtem Food spoils overnight unless blessed or sealed.
\end{enumerate}

\bigskip

%==========================================================
\subsection*{Direwood Encounter Generator}
Roll once when entering a new sector or when the party loses direction.

\paragraph{Step 1: Terrain Distortion (d6)}
\begin{enumerate}
  ℹtem Fog walls−−−movement halved, perception impaired.
  ℹtem Reverse stream−−−water flows upward, whispers of drowned souls.
  ℹtem Forest of empty armor−−−hollow suits clatter as wind passes.
  ℹtem Birch grove−−−trees lean inward as if listening.
  ℹtem Corpse orchard−−−fruit grows from ribcages (\emph{dangerous temptation}).
  ℹtem Silent clearing−−−sound dies, even thoughts echo.
\end{enumerate}

\paragraph{Step 2: Threat Manifestation (d6)}
\begin{enumerate}
  ℹtem Mistwraith Pack hunting fear scent.
  ℹtem Shadewalker patrol marching in silent formation.
  ℹtem Hunger Wolves circling camp perimeter.
  ℹtem Memory echo forcing PCs into historical reenactments.
  ℹtem Voice of the Heroine offering cryptic aid (or condemnation).
  ℹtem \emph{No creature}−−−but the woods rearrange behind them.
\end{enumerate}

\paragraph{Step 3: Psychological Pressure (d6)}
\begin{enumerate}
  ℹtem One PC hears their name whispered from a corpse.
  ℹtem A random PC dreams while awake−−−vision interrupts action.
  ℹtem Someone's equipment rots or tarnishes instantly.
  ℹtem PCs lose track of time−−−night becomes day without transition.
  ℹtem PCs forget a small fact−−−names, directions, purpose.
  ℹtem PCs find something belonging to them−−−but aged or broken.
\end{enumerate}

\paragraph{Step 4: Strange Reward or Revelation (optional d6)}
\begin{enumerate}
  ℹtem A black feather giving +1d to bravery tests.
  ℹtem A shard of gilt metal humming near the heart of the forest.
  ℹtem A whisper of her true name−−−wrong syllables, but directionally useful.
  ℹtem A bone talisman−−−\emph{Ignore fear once, then crack under stress.}
  ℹtem A half−remembered lullaby that soothes undead spirits.
  ℹtem A map that leads in circles unless the party honors the dead.
\end{enumerate}

\textbf{GM Guidance:} Use 1−−2 results per segment. Combine for escalating horror.

\paragraph{Escalation Triggers}
Increase threat when:
\begin{itemize}
  ℹtem PCs desecrate corpses or mock legends.
  ℹtem Someone tries to rationalize the supernatural.
  ℹtem A vision is ignored or laughed off.
\end{itemize}

At high escalation, the Heroine's Phantom appears−−−not as ally, but as judge.

\paragraph{Role of the Heroine in Act II}
She is not a quest−giver−−−she is \emph{the wound}. PCs must:
\begin{itemize}
  ℹtem earn her attention through reverence or boldness,
  ℹtem learn fragments of her forgotten name,
  ℹtem survive her tests.
\end{itemize}

Only by \textbf{speaking her true name} does her attitude shift from hunter to guide.

\subsection*{The Name That Was Forgotten}

The Direwood will not allow the heroine's name to be spoken lightly. Her identity shattered when she died unremembered; the party must \textbf{gather fragments} through visions, echoes, relics, and mercy.

Each revelation provides one \textbf{Name Fragment}. When all are gathered and spoken with intention, her wrath becomes sorrow−−−and the way forward opens.

\bigskip

\begin{center}
\begin{longtable}{cll}
⊤rule
\textbf{d8} & \textbf{Fragment Source} & \textbf{What the PCs Receive} \\
∣rule
1 & Whisper in dream & A syllable spoken by a weeping voice \\
2 & Relic unearthed & Rune−symbol etched in gilt metal \\
3 & Mercy to dead & A spirit forms a phoneme in frost or ash \\
4 & Heroine's phantom & She screams it in rage, misheard or inverted \\
5 & Shadewalker patrol & Captain repeats her battle−cry fragment \\
6 & Shrine remnant & A carved votive holding half her naming−mark \\
7 & Mistwraith hunt & One wraith mimics her dying breath−sound \\
8 & The wood itself & Wind across hollow birches shapes the final tone \\
⊥tomrule
\end{longtable}
\end{center}

\bigskip

\paragraph{Guidance}

\begin{itemize}
  ℹtem \textbf{Fragments are ambiguous:} each is a syllable, rune, or phonetic gesture.
  ℹtem Players must \emph{interpret}−−−wrong assembly draws her fury.
  ℹtem If spoken prematurely, increase the \textbf{Dread Timer } by +1 and attract a spectral assault.
  ℹtem Once all are gathered, speaking the true name pacifies her and reshapes Act IV.
\end{itemize}

\paragraph{Optional Twist}
The final fragment cannot be heard−−−only \emph{remembered}. A PC must sacrifice a cherished memory to speak it aloud.


%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Act III: The Fortress Heart}

The fortress is not built −−− it is \textit{grown}.  
Petrified roots coil into walls, bone−columns hold roofs of fungus stone, and halls pulse faintly as if remembering breath.

Tone: grim revelation + predatory architecture.

\begin{itemize}
  ℹtem \textbf{Shades} recite campaigns that never resolved.
  ℹtem The \textbf{druid's corpse} still hangs impaled by the gilden spear, a font of eternal fog.
  ℹtem The \textbf{nameless heroine} prowls silently −−− testing, haunting, or confronting.
\end{itemize}

\noindent
If the \emph{Dread Timer } is high, she manifests as a lethal nemesis.  
If the PCs earned reverence or curiosity, she delivers riddles, cryptic lore, or half−truth bargains.

\bigskip

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph{Threats of the Fortress}

\begin{itemize}
  ℹtem \textbf{Mistwraith Sentinels:} absorb Fatigue; attack as whispers that drain resolve.
  ℹtem \textbf{Bone Choir Shades:} animate when disturbed, chanting fragments of battle oaths.
  ℹtem \textbf{Shadewalker Wardens:} undead knights commanded by instinct only −−− bodyguards without a master.
  ℹtem \textbf{Grasping Roots:} attempt to root intruders in place; test Body + Athletics to break free.
  ℹtem \textbf{Reflected Self:} mirrors show PCs as future undead −−− risk Mental Strain on failed Resolve.
\end{itemize}

\bigskip

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph{Fortress Traps and Wards}

\textbf{Magical Traps}
\begin{itemize}
  ℹtem \textbf{Bone−Sigil Wards:} trigger phantom spear volleys.
  ℹtem \textbf{Echo Brands:} mark targets −−− nearby undead treat branded PCs as hated traitors.
  ℹtem \textbf{Mist Spore Bursts:} cause hallucinations and misplaced action declarations.
  ℹtem \textbf{Remembrance Locks:} doors open only when spoken with a name fragment.
\end{itemize}

\textbf{Mundane Traps}
\begin{itemize}
  ℹtem Root−pit collapse leading to a crypt nursery of Shadewalkers.
  ℹtem Impaling stakes grown from bone growths under leaf cover.
  ℹtem Hollow floors that drop PCs into memory−echo chambers.
  ℹtem Glyph−etched bells that summon a Shade patrol if rung.
\end{itemize}

\bigskip

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph{Room Generator: The Dire Stronghold}

Roll d10 for each chamber or turn deeper:

\begin{center}
\begin{longtable}{cl}
⊤rule
\textbf{d10} & \textbf{Room / Encounter} \\
∣rule
1 & Hall of Hollow Shields −−− spectral infantry drill endlessly \\
2 & Bone Spindle Stair −−− roots twist upward; false steps activate impaling branches \\
3 & The Echo Choir −−− Shades lament their failed charge; listening risks Fatigue \\
4 & The Memory Pit −−− visions drag one PC into a relived massacre \\
5 & Oath Archive −−− bone tablets with names scored out; fragments hidden here \\
6 & The Bloodless Barracks −−− Shadewalker knights frozen until disturbed \\
7 & Grove of Still Hearts −−− petrified Druid seeds containing wraith embryos \\
8 & Hall of the Betrayed −−− mirrors show the PCs as corrupted avatars \\
9 & Root−Chamber Gate −−− door opens only to a true fragment spoken aloud \\
10 & The Spear Hall −−− the druid's husk nailed to reality; the heroine watches \\
⊥tomrule
\end{longtable}
\end{center}

\bigskip

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph{Adversary Table: Fortress Foes}

\begin{center}
\begin{longtable}{cl}
⊤rule
\textbf{d6} & \textbf{Adversary} \\
∣rule
1 & Shade Cohort (3−−5 spectral infantry, DV 3 vs Resolve) \\
2 & Mistwraith Pack (absorbs Fatigue before Harm) \\
3 & Shadewalker Knight (Cap 3 undead bodyguard wielding rusted rites) \\
4 & Echo Sorcerer (fragment of the druid, casts DV 4 control spells) \\
5 & The Forgotten Squire (undead child bearing a name rune fragment) \\
6 & Her Shrine−Guardian Self −−− a phantom of the heroine in life \\
⊥tomrule
\end{longtable}
\end{center}

\bigskip

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph{Environmental Effects (1d6 per room)}

\begin{itemize}
  ℹtem \textbf{1: Breath Leech} −−− PCs lose 1 Fatigue entering.
  ℹtem \textbf{2: Time Loop} −−− first action each turn repeats.
  ℹtem \textbf{3: Whispering Winds} −−− disadvantage on stealth.
  ℹtem \textbf{4: Grief Pulse} −−− one PC makes a sorrow check.
  ℹtem \textbf{5: Bone Bloom} −−− undead roots sprout mid−combat.
  ℹtem \textbf{6: Heroine's Watch} −−− her phantom judges the action; aid or hinder.
\end{itemize}

\bigskip

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph{GM Guidance: The Heroine as Dynamic Antagonist}

She is not merely a boss −−− she is a \emph{mirror}.

\begin{itemize}
  ℹtem When PCs show \textbf{oath−keeping}, she grants a name fragment.
  ℹtem When they show \textbf{hubris or desecration}, treat her as a deadly hunter.
  ℹtem If her name is assembled, the fortress reorganises −−−  
  root−walls peel back, the mist recoils, and the druid's corpse becomes vulnerable.
\end{itemize}

\noindent
This act should end with a revelation, not a combat −−− unless the party refuses to remember.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

%====================================================
\subsection{Name Quest Subsystem: Remembering the Forgotten}
łabel{subsec:name_quest}
∈dex{Name Quests}

\paragraph*{Premise}
Some truths cannot be killed −−− only forgotten.  
The Heroine of the Mist can only be freed by remembering her.

⫕section*{Memory Track}
A shared party track with 6 boxes.

Mark one when:
\begin{itemize}
    ℹtem uncovering a name fragment or true deed,
    ℹtem fulfilling an oath tied to her story,
    ℹtem suffering hardship to defend the innocent,
    ℹtem being recognized by shades as her lost companions.
\end{itemize}

\paragraph*{Effects at Thresholds}
\begin{description}
    ℹtem[2 Marks:] Shades hesitate; mist retreats briefly.
    ℹtem[4 Marks:] The Heroine shifts from Nemesis to Trial.
    ℹtem[6 Marks:] PCs may speak the Name −−− final resolution triggers.
\end{description}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Name Fragments}
Fragments may be discovered, traded, stolen, or revealed through:
\begin{itemize}
    ℹtem prophetic dreams,
    ℹtem battlefield memories,
    ℹtem ghost−songs,
    ℹtem the Golden Thorn's revelations.
\end{itemize}

Each Name Fragment must be \textbf{correctly placed} in her saga −−− requiring Roleplay + Lore + Presence.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Final Rite: Restoration of Name}
When all fragments fit:
\begin{description}
    ℹtem[Her Shade kneels,] ceasing hostility.
    ℹtem[The Mist lashes,] then withdraws.
    ℹtem[The Heroine ascends,] becoming patron or guide.
\end{description}

The party must swear:
\textit{her saga shall be sung,  
her deeds recorded,  
her name remembered.}

Breaking this oath:
\begin{itemize}
    ℹtem summons her wrath,
    ℹtem marks all PCs as \textbf{oathbreakers},
    ℹtem the Mist becomes a nemesis for the rest of the campaign.
\end{itemize}

%====================================================
%====================================================
⫕section*{Act IV: Descent into the Root−Vault}

Tone: metaphysical pursuit + sacrificial myth

\paragraph{Premise}
The Root−Vault is an inverted under−world where:
\begin{itemize}
  ℹtem roots are ossified memory,
  ℹtem words become walls,
  ℹtem and oaths have gravity.
\end{itemize}

Here, sealed at the breach−point of the First Curse,
rests the \textbf{Resonant Keystone} −−−  
a crystalline bell−core wrapped in the heroine's erasure.

\paragraph{Why It Matters}
Without it, the bell−line can never be repaired −−−
the Mist spreads until shore cities fall.

With it, PCs can:
\begin{itemize}
  ℹtem re−tune the warding line,
  ℹtem awaken the quiet bells,
  ℹtem and name the heroine back into existence.
\end{itemize}

\paragraph{Vault Threats}

\begin{itemize}
  ℹtem \textbf{Name−Eaters} −−−  
        spectral archivists that rip identity pages from memory.

  ℹtem \textbf{Mist Stalkers} −−−  
        ragged emanations of the druid's will,
        trying to seize the Keystone to unbind the curse.

  ℹtem \textbf{Root−Engines} −−−  
        biomechanical heart nodes that reorganize corridors,  
        threatening to erase the path behind PCs.
\end{itemize}

\paragraph{Trial of Worth}  
To claim the Keystone, PCs must:

\begin{enumerate}
  ℹtem \textbf{Speak or fully assemble the heroine's name.}
  ℹtem \textbf{Acknowledge her sacrifice −−− and its injustice.}
  ℹtem \textbf{Swear an oath to restore her saga publicly.}
\end{enumerate}

Oath refusal?  
The Keystone remains inert −−− or lashes out violently.

\paragraph{Backlash Possibilities}
If PCs try to take the Keystone without the oath:
\begin{itemize}
  ℹtem the heroine becomes their eternal Nemesis,
  ℹtem the Resonant Keystone fractures (marked with their names),
  ℹtem and Act V begins with Payden's Port collapsing.
\end{itemize}

\paragraph{Act IV Ends When:}

\begin{itemize}
  ℹtem the PCs physically hold the \textbf{Resonant Keystone}, and
  ℹtem their bond −−− or enmity −−− with the heroine is sealed.
\end{itemize}
%====================================================

%====================================================
\subsection*{Act V: The Last Saga −−− The Seal and the Spear}
łabel{act5:lastₛaga}

Tone: mythic horror, tragic revelation, ritual triumph

\noindent
The Heart−Fortress shudders as roots split, mist curls in reverse,  
and the pinned druid's body flexes against the Golden Thorn.  
The party stands where the curse began and will end −−− or repeat.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{The Awakening of the Husk}

If the PCs disturbed the Thorn, or if the Name remains incomplete:

\begin{itemize}
  ℹtem The \textbf{Druid of the Mist} reanimates.
  ℹtem Root−bone walls shatter into a \textbf{spiral arena}.
  ℹtem The husk becomes a \textbf{multi−phasic boss} (see Stat Block).
\end{itemize}

The party cannot kill it −−− only \emph{subdue} or \emph{pin} it again.  
Its flesh reforms, its bones knit; it is the curse given shape.

\paragraph*{Final Objective:}  
Hold long enough to force it beneath the Thorn.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Intervention of the Heroine Shade}

If the PCs completed the Name Quest (Section~\ref{subsec:name_quest}):

\begin{itemize}
  ℹtem The Heroine Shade manifests in full Valkyric radiance.
  ℹtem She strides through the mist untouched, spear alight.
  ℹtem Her presence weakens the husk's regeneration.
\end{itemize}

\paragraph{Resolution Trigger}
A PC must proclaim her Name aloud as the Shade strikes −−−  
only then does the cycle break.

\medskip
If done, the Shade impales the Druid again with:
\begin{itemize}
  ℹtem golden light,
  ℹtem cascading roots,
  ℹtem and the sighs of a thousand dead.
\end{itemize}

The arena collapses inward, and the mist recoils violently.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{If the Name Was Not Restored}

The husk overwhelms the fortress's structure:

\begin{itemize}
  ℹtem Root corridors twist and collapse.
  ℹtem Mist seals passages behind the party.
  ℹtem They must flee through the \textbf{Root−Vault escape path}.
\end{itemize}

Survival is victory −−− but the curse remains, stronger than before.  
The Golden Thorn becomes inert until her Name is remembered in future play.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Loot, Legacy, and Consequences}

\paragraph{If the Name Was Restored}
\begin{itemize}
  ℹtem PCs gain \textbf{Saga Reputation +1}.
  ℹtem The Heroine becomes a \textbf{Patron of Valor, Memory, and Resurrection}.
  ℹtem The Mist withdraws for a generation −−− bell−lines strengthen.
  ℹtem The Golden Thorn ascends in power (see Artifact Upgrades).
\end{itemize}

\paragraph{If the PCs Only Re−Pinned the Druid}
\begin{itemize}
  ℹtem The fortress stabilises.
  ℹtem The Mist is contained, not ended.
  ℹtem PCs become \textit{known} −−− but without honor or clarity.
  ℹtem The Thorn is unchanged, awaiting remembrance.
\end{itemize}

\paragraph{If the PCs Failed or Fled}
\begin{itemize}
  ℹtem Mist incursions increase for months.
  ℹtem Bell−lines must be rebuilt at great cost.
  ℹtem PCs are viewed with suspicion −−− or blame.
\end{itemize}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Epilogue: The Heroine's Choice}

If restored:
\begin{quote}
``My saga is yours −−− now yours is mine.
Go −−− and let the living remember.''
\end{quote}

\noindent
She dissolves into starlight −−− but may return as:
\begin{itemize}
  ℹtem a Patron,
  ℹtem a guide spirit,
  ℹtem or a PC boon (once/arc miracle of defense or intervention).
\end{itemize}

If not restored, her Shade remains bound:
\begin{itemize}
  ℹtem sometimes an ally,
  ℹtem more often a Nemesis in later arcs,
  ℹtem forever seeking one who will speak her name.
\end{itemize}

%====================================================

\paragraph{GM Guidance: Multi−Phasic Boss Confrontation}

Removing the tuning−spike or attempting to destroy the husk awakens the \textbf{First Druid} in a catastrophic resurrection state.  
This is not a standard fight −−− it is a ritualized horror battle with shifting phases and survival focus.

\subparagraph{Phase I: The Husk Wakes}
\begin{itemize}
  ℹtem The petrified body reanimates, still impaled.
  ℹtem It lashes mist tendrils and summons \emph{root−shade guardians}.
  ℹtem Attacks ignore armor unless the PCs invoke sacred names or oaths.
\end{itemize}

\noindent % spacing
\textbf{Objective:} Break its stabilization −−− the spike needs to be removed or driven \emph{deeper}.

\subparagraph{Phase II: The Unbound Druid}
When the spike shifts:
\begin{itemize}
  ℹtem The husk splits open along its old wounds.
  ℹtem The druid's true form −−− a writhing tangle of bone, antler, and liquid mist −−− pulls free.
  ℹtem Mist thickens; PCs must navigate collapsing terrain and hallucinations.
\end{itemize}

\noindent
\textbf{Abilities:}
\begin{itemize}
  ℹtem \emph{Corrupted Command} −−− rerolls PC successes into failures in scenes of desecration.
  ℹtem \emph{Mist Rebirth} −−− dismissed enemies reform unless sanctified.
\end{itemize}

\noindent
\textbf{Objective:} Force the druid back to the sanctum−heart −−− where the spike can restrain it again.

\subparagraph{Phase III: The Heart Rift}
Push the druid to the spike altar −−− the battlefield becomes a circular pit of oaths, death echoes, and sagas written in air.

\begin{itemize}
  ℹtem PCs must endure mist storms, identity erosion, and memory theft.
  ℹtem The druid attempts to \emph{rewrite one PC into its replacement host}.
\end{itemize}

\noindent
\textbf{Objective:} Impale the druid again −−− but only through a sacrifice, bargain, or declared deed.

\subparagraph{Conditional Intervention: The Heroinic Shade}
If the party has restored the heroine's name:
\begin{itemize}
  ℹtem Her shade manifests behind the druid as a golden silhouette.
  ℹtem She seizes the spike herself and drives it home through the druid's heart−core.
  ℹtem Mist recoils; servants disintegrate; the sanctum stabilizes.
\end{itemize}

\noindent
This ending:
\begin{itemize}
  ℹtem restores the status quo,
  ℹtem seals the husk back into slumber,
  ℹtem and confirms the heroine as a protector rather than nemesis.
\end{itemize}

\subparagraph{If Her Name Is Not Restored}
The fight must continue through three additional rounds of escalating horror:

\begin{itemize}
  ℹtem the druid's antler mass becomes a crown of thorns,
  ℹtem a PC must willingly offer an oath and blood to bind the spike,
  ℹtem otherwise the fortress collapses and the Mist erupts outward.
\end{itemize}

\noindent
If the fight ends without the heroine intervening, the victory is pyrrhic −−−  
the druid is bound, but the Mist threshold expands, warping the wider region.

\subparagraph{Reward for the True Name Path}
If her true saga is spoken before or during the battle:
\begin{itemize}
  ℹtem the fight ends early in dramatic fashion,
  ℹtem the heroine becomes a continuing NPC patron or ally,
  ℹtem PCs may take her boon or blessing as a campaign−long tag.
\end{itemize}

\noindent
This version preserves high tension while allowing the name−restoration path to feel mythic and consequential.

\paragraph{Act III Ends When:}
\begin{itemize}
  ℹtem PCs understand that true salvation lies below,
  ℹtem their relationship to the heroine is set (Nemesis, Test, or Maybe Ally),
  ℹtem and the path to the vault opens −−− willingly or through collapse.
\end{itemize}
%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−

%====================================================
\subsection{Key Threats}

⫕section*{Mistwraiths (Tier IV Horrors)}

Secrets made flesh. Feed on awareness and fear.  
Appear where the party hesitates or falters.

⫕section*{Shadewalkers (Tier III Undead)}

Corpse−puppets bound in root and bone.  
Slow, relentless, territorial.

⫕section*{Direwood Beasts}

Animals twisted by necromantic saturation.  
Symptoms: extra limbs, eyeless faces, echoing screams.

⫕section*{The Nameless Heroine}

Shade−queen, tragic antagonist.  
If encountered before her name is found, treat her as a boss encounter.  
If her name is revealed mid−battle, she collapses into grief rather than rage.

%====================================================
\subsection{Environmental Horror}

\paragraph{Core Tags:}
\begin{itemize}
  ℹtem \textbf{[MIST−CHOKED]} −−− Disorienting and oppressive.
  ℹtem \textbf{[HAUNTED]} −−− Reality shows seams; time loops.
  ℹtem \textbf{[STARVED−WORLD]} −−− No food, no warmth, whispers that sap Resolve.
\end{itemize}

Consider starvation, dehydration, and will−despair as slow pressures.  
The Direwood wants them to give up−−−or accept its bargain.

%====================================================
\subsection{Rewards and Consequences}

⫕section*{Curse−Breaker Outcomes}

\begin{itemize}
  ℹtem \textbf{Saga−Bearer:} PCs gain renown as keepers of a forgotten truth.
  ℹtem \textbf{Artifact Wardens:} They inherit the gilden spear−−−now dormant unless wielded in her name.
  ℹtem \textbf{Map Change:} The Mistline shifts; local politics scramble to react.
\end{itemize}

⫕section*{Failure or Refusal}

\begin{itemize}
  ℹtem The Mist expands or mutates.
  ℹtem The heroine becomes a recurring nemesis.
  ℹtem PCs are marked by the Direwood−−−hunted by nightmares or undead emissaries.
\end{itemize}

%====================================================
\subsection{Using With Any Setting}

Plug this adventure into:

\begin{itemize}
  ℹtem a Viking−flavored frontier,
  ℹtem a Slavic cursed forest,
  ℹtem a Romanized empire borderland,
  ℹtem a gothic or Celtic wilderness,
  ℹtem or a colonial frontier horror setting.
\end{itemize}

Swap religions, factions, and cultural icons as needed−−−the core dynamics remain:

\begin{itemize}
  ℹtem \textbf{Forgotten hero as antagonist}
  ℹtem \textbf{Curse anchored in erasure of identity}
  ℹtem \textbf{The truth is the weapon, not the sword}
\end{itemize}

This keeps the module stand−alone, portable, and emotionally powerful.

\begin{center}
    {Łarge\bfseries Tales of the Direwood}\\[4pt]
    \textit{A Traveler's Briefing for Those Who Walk Beyond the Bell−Line}
    \end{center}
    
    \vspace{0.5em}
    
    \noindent
    \textbf{What is the Direwood?}\\
    A blighted forest where fog is older than the trees.\
    No map draws the same twice.\
    Beasts walk on borrowed bones.\
    Time forgets to move correctly.
    
    \vspace{0.6em}
    
    \noindent
    \textbf{What the Elders Say:}
    \begin{itemize}
        ℹtem Once it guarded the borderlands.
        ℹtem Now it breathes death instead.
        ℹtem Something within remembers us.
    \end{itemize}
    
    \vspace{0.4em}
    
    \noindent
    \textbf{Safe Conduct Within the Mist}
    \begin{itemize}
        ℹtem Travel in pairs; loneliness draws attention.
        ℹtem Firelight helps, but only if you believe it will.
        ℹtem Bells keep the lesser spirits away−−−until they learn you.
        ℹtem Do not chase echoes; if someone calls your name, check if their feet touch the ground.
    \end{itemize}
    
    \vspace{0.4em}
    
    \noindent
    \textbf{Signs of Corruption}
    \begin{itemize}
        ℹtem Roots that bleed when cut.
        ℹtem Faces where bark should be.
        ℹtem Whispers in languages you almost know.
    \end{itemize}
    
    \vspace{0.6em}
    
    \noindent
    \textbf{Things Seen Near the Heartwood}
    \begin{itemize}
        ℹtem Towers of petrified wood and bone.
        ℹtem A spearhead of gold hammered through stone.
        ℹtem A phantom warrior in ruined mail watching from the mist.
    \end{itemize}
    
    \vspace{0.6em}
    
    \noindent
    \textbf{The Nameless Hero}
    \begin{itemize}
        ℹtem Locals speak of the "Warden Without Name."
        ℹtem Some call her guardian, others curse her as betrayer.
        ℹtem Legend holds she fought a great evil−−−and died forgotten.
    \end{itemize}
    
    \vspace{0.6em}
    
    \noindent
    \textbf{Rumors and Warnings}
    \begin{itemize}
        ℹtem \textit{"If you speak her name, she might listen."}
        ℹtem \textit{"If you mock her, she will find you."}
        ℹtem \textit{"The spear keeps him dead−−−but keeps her bound."}
    \end{itemize}
    
    \vspace{0.6em}
    
    \noindent
    \textbf{Why Venture In?}
    \begin{itemize}
        ℹtem To stop the creeping Mist.
        ℹtem To find lost knowledge.
        ℹtem To free or defeat the Warden.
        ℹtem To discover her true name.
    \end{itemize}
    
    \vspace{0.8em}
    
    \noindent
    \textbf{Advice from Survivors}
    \begin{itemize}
        ℹtem Bring rope, salt, bells, and flame.
        ℹtem Do not trust your memories inside.
        ℹtem When in doubt−−−speak truth, not courage.
    \end{itemize}
    
    \vspace{1.0em}
    
    \noindent
    \begin{center}
    \textit{The Direwood does not hate you.}\\
    \textit{It simply remembers before you do.}
    \end{center}

    %====================================================
⫕section{Rites of the Druid of the Mist}
łabel{rites:mistrdruid}
∈dex{Rites!Druid of the Mist}

\noindent
These rites reflect the patron's ethos: \emph{erode, unmake, devour, dissolve}.

Followers learn that every invocation is a transaction −−− power traded for identity.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph*{Low Rites (4−−6 XP)}

\begin{description}

  ℹtem[Mist's Caress] \emph{Scene; Zone; No.} \\
  \texttt{[VEIL] [SENSE] [DRAIN]} \\
  Call thin fog across Near. Gain +1 die to detect fear, fatigue, or spiritual imbalance.  
  Living foes suffer unease; undead gain interest.

  ℹtem[Whisper of Hollow Roots] \emph{Scene; Self; No.} \\
  \texttt{[SENSE] [GUIDE]} \\
  Hear paths where boundaries blur. Once per scene, bypass mundane impediment (wall, guard, protocol) narratively; GM provides complication.

  ℹtem[Sap of Forgetting] \emph{Scene; Touch; Yes.} \\
  \texttt{[DRAIN] [WEAKEN]} \\
  Inflict \(−1\) die to resistance rolls for one exchange as the target's name wavers.  
  Push It: Target forgets one small truth or detail temporarily.

  ℹtem[Bitter Blossom Mark] \emph{Extended; Close; Yes.} \\
  \texttt{[CURSE] [DECAY]} \\
  Plant a sigil that corrupts soil, masonry, or a social oath.  
  Start a \textbf{Creeping Ruin [4]} timer; when full, something breaks.
  
\end{description}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph*{Standard Rites (7−−9 XP)}

\begin{description}

  ℹtem[Root of Unmaking] \emph{Scene; Zone; Yes.} \\
  \texttt{[DECAY] [BIND] [CORRUPT]} \\
  Infect a structure, body, or agreement with invisible rot.  
  Allies gain +1 die to exploit fractures; order−based actions suffer +1 DV.

  ℹtem[Shroud of Dissolution] \emph{Scene; Self; Yes.} \\
  \texttt{[VEIL] [CONCEAL]} \\
  Become indistinct: sound muffles, form blurs. Gain +1 die on evasion/stealth; fail with 1s risks name−fragment loss (GM chooses).

  ℹtem[Bone Orchard Bloom] \emph{Scene; Zone; Yes.} \\
  \texttt{[ANIMATE] [AREA] [CORRUPT]} \\
  Corpses or dead roots animate as spectral vines/limbs that hinder foes.  
  Create \textbf{Difficult Terrain} and start a \textbf{Gravegarden [4]} timer.

  ℹtem[Harvester's Breath] \emph{Scene; Zone; Yes.} \\
  \texttt{[DRAIN] [FEED] [AREA]} \\
  Draw fatigue, sorrow, or fear from all in Near.  
  Convert one such condition into +1 die for yourself or a chosen ally.

\end{description}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph*{High Rites (10−−14 XP)}

\begin{description}

  ℹtem[Reclaim the Boundary] \emph{Scene; Zone; Yes.} \\
  \texttt{[DOMAIN] [CURSE] [REWRITE]} \\
  Select a place or law; for the scene, its borders cease functioning.  
  Walls don"t divide, ranks don"t compel, names are unreliable.  
  Resistance causes sensory fragmentation (Position −1 next exchange).

  ℹtem[Pale Menagerie] \emph{Scene; Near; Yes.} \\
  \texttt{[SUMMON] [UNDEAD] [COMMAND]} \\
  Call forth 1−−3 mist−echoes (Cap 2) that obey simple urges.  
  They dissolve when named or sanctified.

  ℹtem[Walk of Unbeing] \emph{Scene; Self; Yes.} \\
  \texttt{[PHASE] [VEIL] [MOVE]} \\
  Step through boundaries, walls, or social constraints.  
  Treat any Blocked route as Open but mark \textbf{1 Name Erosion}.

  ℹtem[Famine of the Living Word] \emph{Scene; Far; Yes.} \\
  \texttt{[CURSE] [SILENCE] [BLIGHT]} \\
  Silence a leader, liturgy, or institution.  
  For one scene, oaths, commands, or blessings falter.  
  GM marks \textbf{Narrative Ruin} −−− consequences linger.

\end{description}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
\paragraph*{Apex Rite (15−−20 XP)}

\begin{description}

  ℹtem[The Mist Devours the Story] \emph{Extended; Domain; Yes.} \\
  \texttt{[UNMAKE] [DOMAIN] [APOCALYPSE]} \\
  Declare a rooted structure −−− a keep, dynasty, covenant, religion, or name.  
  Begin \textbf{The Unmaking [6+]} timer.  
  When full, that thing loses identity:  
  histories rewrite, memories fracture, its power−base dissolves.  
  Survivors may cling to fragments, but it never becomes the same again.  
  \emph{Cost:} Mark \textbf{3 Erosion}, lose one personal truth, risk becoming an avatar.

\end{description}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Corruption & Erosion Mechanics}
∈dex{Corruption}
∈dex{Identity Loss}

Followers of the Mist do not track \emph{Obligation}; they track \textbf{Erosion}.

Each time a rite produces backlash or a 1 is rolled:
\begin{itemize}
  ℹtem erase a memory,
  ℹtem distort a feature,
  ℹtem or lose a social bond.
\end{itemize}

At \textbf{3+ Erosion}, the GM may enforce:
\begin{itemize}
  ℹtem shadow−speech,
  ℹtem hostile mist manifestations,
  ℹtem or NPCs forgetting the PC's name.
\end{itemize}

At \textbf{6+ Erosion}, the patron attempts apotheosis −−−  
the character becomes a \emph{Mist−Thing} unless named and anchored by others.

%====================================================

%====================================================
⫕section{Spellwheel: The Druid of the Mist}
łabel{spellwheel:mistrdruid}

\begin{center}
\begin{longtable}{|c|c|c|p{6cm}|}
ℎline
\textbf{Sigil} & \textbf{Aspect} & \textbf{Tags} & \textbf{Invocation Theme} \\
ℎline
Root & Boundary & [BIND][DECAY] &
Unmake structure, rot identity, bind paths or truths \\
ℎline
Bloom & Hunger & [DRAIN][FEED] &
Harvest fear/fatigue; convert suffering into momentum \\
ℎline
Mist & Veil & [VEIL][SENSE] &
Cloud meaning; detect cracks, sorrow, forgotten names \\
ℎline
Bone & Unlife & [ANIMATE][CORRUPT] &
Raise echoes; twist flesh or soil with old death \\
ℎline
Hollow & Dominion & [CURSE][REWRITE] &
bend will, rewrite laws, fracture names and memories \\
ℎline
Rift & Tide of Unbeing & [PHASE][UNMAKE] &
erase borders, dissolve authority, or step through reality \\
ℎline
\end{longtable}
\end{center}

\paragraph*{Usage Notes}
Cantors assemble invocations by chaining:
\[
\textit{Aspect} + \textit{Tags} + \textit{Channel}
\]
Invoker−style play treats the wheel as a \emph{menu}:
\begin{enumerate}
    ℹtem Choose 1−−2 Aspects
    ℹtem Add 1−−3 matching Tags
    ℹtem State the narrative form (fog, vines, sorrow, whispers)
\end{enumerate}

\noindent
The more the invocation erodes boundaries or identity,  
the higher the DV and the greater the Erosion risk.

%====================================================

%====================================================
\subsection{Patron: The Heroine of the Mist}
łabel{patron:mistₕeroine}
∈dex{Patrons!Heroine of the Mist}

\paragraph*{Aspect}
\textit{Memory, Oath, Bloodshed in Defense}

Once mortal −−− now half−Valkyrie, half−Mourn−Warden.  
She embodies:
\begin{itemize}
    ℹtem valor without recognition,
    ℹtem sacrifice without audience,
    ℹtem and the terrible cost of defending those who never remember you.
\end{itemize}

\paragraph*{Sphere of Influence}
\begin{itemize}
    ℹtem names,
    ℹtem sagas,
    ℹtem identity persistence,
    ℹtem righteous vengeance,
    ℹtem guardianship of the innocent,
    ℹtem oaths that bind but uplift.
\end{itemize}

\paragraph*{Devotional Archetypes}
\begin{description}
    ℹtem[Oath−Keepers:] sworn blades, bodyguards, shieldbearers
    ℹtem[Saga−Wardens:] skalds, lorekeepers, the memory−obsessed
    ℹtem[The Forgotten:] outcasts, orphans, traumatised veterans
\end{description}

\paragraph*{Taboos}
\begin{itemize}
    ℹtem abandoning a sworn charge,
    ℹtem erasing someone's rightful name,
    ℹtem cowardice in duty.
\end{itemize}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section{Miracles and Major Rites}
Devotional power flows through remembrance and sacrifice.

\paragraph*{Rites (General Access)}
\begin{description}
    ℹtem[Namebinding:] Mark 1 Fatigue −−− grant an ally \textbf{+1 Position} for defending or reclaiming identity.
    ℹtem[Saga's Spark:] Once per scene, declare a past victory to gain \textbf{+1 die} for an action −−− must narrate the remembered deed.
    ℹtem[Shield of the Forgotten:] Allies under your protection count as \textbf{one tier higher} for resisting fear or domination.
    ℹtem[Blood Oath:] Swear protection over a person or truth; gain a boon whenever fulfilling that oath through hardship.
    ℹtem[Witness to Glory:] Convert suffering (Fatigue/Harm) into +1 SB when defending others.
\end{description}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section{Higher Mysteries (Prestige/Epic Rites)}
\begin{description}
    ℹtem[Reclaim the Name:] Restore a forgotten truth or identity; remove Mist penalties for the scene.
    ℹtem[Mist−Splitting Spear:] Manifest spectral spear; +2 dice vs undead and corrupted oathbreakers.
    ℹtem[Saga Ascent:] Once per session, prevent death of one ally −−− they rise with 1 Harm and a memory of you they cannot shake.
    ℹtem[The Last Watch:] Bind your spirit to a place or charge; if abandoned, you become a shade that hunts oathbreakers.
\end{description}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section{Spellwheel: The Heroine of the Mist}
łabel{spellwheel:mistₕeroine}

\begin{center}
\begin{longtable}{|c|c|c|p{6cm}|}
ℎline
\textbf{Sigil} & \textbf{Aspect} & \textbf{Tags} & \textbf{Invocation Theme} \\
ℎline
Spear & Valor & [SMITE][PURIFY] & punish undead, strike corruption, reveal truth \\
ℎline
Shield & Oath & [GUARD][ENDURE] & intercept harm, stand immovable, sanctify ground \\
ℎline
Torch & Memory & [REVEAL][NAME] & expose hidden identity; illuminate forgotten stories \\
ℎline
Tears & Sacrifice & [BIND][HEAL] & heal through cost; bind wounds, take suffering \\
ℎline
Crown & Remembrance & [SAGA][INSPIRE] & uplift allies, declare deeds into reality, rewrite legacy \\
ℎline
Bell & Requiem & [BANISH][SEAL] & end hauntings; sever curses; close breaches \\
ℎline
\end{longtable}
\end{center}

\paragraph*{Cantor/Invoker Themes}
\begin{itemize}
    ℹtem songs of valor,
    ℹtem whispered sagas,
    ℹtem spear−light metaphors,
    ℹtem radiant mist,
    ℹtem oath−fire halos.
\end{itemize}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section{Devotional Gifts (Mechanical Boons)}
\begin{itemize}
    ℹtem \textbf{Once per session:} treat any Desperate defense as Dominant.
    ℹtem When defending others, gain \textbf{+1d} to resist fear, torment, or loss.
    ℹtem When fulfilling sworn duty under hardship, \textbf{mark 1 SB}.
\end{itemize}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section{Narrative Role in the Direwood Adventure}
If her true Name is restored:
\begin{itemize}
    ℹtem she manifests at Act V,
    ℹtem impales the Druid with the \textit{Golden Thorn},
    ℹtem and offers one PC her blessing −−− or burden.
\end{itemize}

If her Name remains lost:
\begin{itemize}
    ℹtem she remains a Nemesis Shade,
    ℹtem hunting oathbreakers and defilers,
    ℹtem until someone speaks her back into the world.
\end{itemize}

%====================================================

%====================================================
\subsection{Nemesis Profile: The Heroine Shade}
łabel{nemesis:mistₕeroine}
∈dex{Nemesis!Heroine of the Mist}

\paragraph*{Role}
\textit{Guardian of memory, judge of oaths, hunter of the unworthy.}

A spectral shieldmaiden formed of mist and radiant sorrow.  
She is tragedy weaponised.

\paragraph*{Tier}
IV (Legendary Adversary)

\paragraph*{Concept Tags}
[VALKYRIE] [OATH] [MOURN] [JUDGE] [SAGA]

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Core Traits}
\begin{description}
    ℹtem[Spirit:] 6
    ℹtem[Resolve:] 5
    ℹtem[Body:] 4
    ℹtem[Speed:] 3
    ℹtem[Presence:] 6
\end{description}

\noindent
\textbf{Harm Capacity:} 6 (Cannot be killed until her Name is restored)

\textbf{Armor:} 2 (Spectral Plate)

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Phase Structure}

\paragraph*{Phase I: Silent Hunt}
\textit{She watches, tests, strikes symbols not flesh.}

\begin{itemize}
    ℹtem Attacks inflict Conditions rather than Harm:\\
    [REVERENCE], [FEAR], [UNWORTHY]
    ℹtem If PCs perform \textbf{oath−keeping}, she withdraws for 1 exchange.
\end{itemize}

\paragraph*{Phase II: The Shieldmaiden Revealed}
\textit{Triggered when she is struck for 3+ Harm in one scene.}

\begin{itemize}
    ℹtem Gains +2 dice to melee attacks.
    ℹtem Her spear manifests: Harm 2, ignores Armor.
    ℹtem She utters fragments of her lost Name.
\end{itemize}

\paragraph*{Phase III: The Ascendant Shade}
\textit{Triggered when PCs hold three Name Fragments or swear an oath to restore her saga.}

\begin{itemize}
    ℹtem She switches from Nemesis to Trial.
    ℹtem PCs may \textbf{roll to remember} (Lore + Presence).
    ℹtem If spoken true, she kneels −−− and joins Act V.
\end{itemize}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Abilities}

\paragraph*{Mist−Splitting Spear}
+2 dice vs corrupted or undead foes; on 6s, pins enemy spirits.

\paragraph*{Witness of Deeds}
Once per scene, invert Position against oathbreakers.

\paragraph*{Saga Binding}
When PCs narrate past deeds tied to sacrifice, grant +1d to them.

\paragraph*{Oathscour}
When someone abandons a charge, immediately strike Harm 2 ignoring Armor.

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Weaknesses}
\begin{itemize}
    ℹtem Cannot be harmed until her Name is known.
    ℹtem Will not strike the innocent unless manipulated.
    ℹtem Spear shattered by betrayal or false remembrance.
\end{itemize}

%====================================================

%====================================================
\subsection{Artifact: The Golden Thorn}
łabel{artifact:goldenₜhorn}
∈dex{Artifacts!Golden Thorn}

\paragraph*{Origin}
Forged by an unknown power to pierce immortal spirits and bind mist−born corruption.

\paragraph*{Appearance}
A meter−long gilded pin, root−etched, warm to the touch, humming faintly with invoked names.

\paragraph*{Properties}
\begin{itemize}
    ℹtem \textbf{Harm 2;} ignores Armor on undead or cursed entities.
    ℹtem When used to impale, creates a \textbf{Banishment Seal}.
    ℹtem If wielded by one who knows the Heroine's true Name, treat as \textbf{Harm 3}.
\end{itemize}

%−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−−
⫕section*{Golden Thorn Invocation Rites}
These may be used by Cantors, Invokers, or Runekeepers.

\paragraph*{Thorn Seal}
\texttt{[BANISH][SEAL][MIST]} −−− impale a spirit, pinning it to reality.

\paragraph*{Memory Piercing}
\texttt{[NAME][REVEAL]} −−− the impaled target utters a name or truth unwillingly.

\paragraph*{Vigilant Thorn}
\texttt{[GUARD][WARD]} −−− mark a threshold; undead must test to cross.

\paragraph*{Radiant Impalement}
\texttt{[PURIFY][LIGHT][SPEAR]} −−− destroy spectral extensions or illusions.

\noindent
\textbf{Restriction:}  
May only be used at full potency by one who has sworn an oath to restore the forgotten.

%====================================================
\appendix
\chapter{Running Direwood: GM Tips and Appendix}

\section{Preparation Checklist}

\subsection{Session Zero Considerations}
\begin{itemize}
    ℹtem Confirm player comfort with horror themes (psychological horror, identity loss, existential dread)
    ℹtem Establish safety tools (X−Card, Script Change) for intense scenes
    ℹtem Review character backgrounds for integration hooks
    ℹtem Set expectations for investigative vs. combat focus
\end{itemize}

\subsection{Pre−Session Preparation}
\begin{enumerate}
    ℹtem Map key faction NPCs with clear motivations
    ℹtem Prepare 3−4 name fragment discovery scenarios
    ℹtem Review Patron mechanics for The Heroine and The Druid
    ℹtem Set up tracking sheets for:
        \begin{itemize}
            ℹtem Dread Timer 
            ℹtem Memory Track
            ℹtem Faction attitudes
            ℹtem Environmental consequences
        \end{itemize}
    ℹtem Prepare key location descriptions (Payden's Port, Direwood sectors, Fortress Heart)
\end{enumerate}

\section{Pacing and Flow Management}

\subsection{Act I: Payden's Port}
\begin{itemize}
    ℹtem \textbf{Focus}: Political tension + creeping horror
    ℹtem \textbf{Key Scenes}: Faction meetings, first breach, departure
    ℹtem \textbf{Pacing}: 2−3 sessions
    ℹtem \textbf{GM Tips}: 
        \begin{itemize}
            ℹtem Use omens liberally (dead gulls, fog children)
            ℹtem Let factions compete for PC attention
            ℹtem Don't rush the departure − build dread
        \end{itemize}
\end{itemize}

\subsection{Act II: Through the Direwood}
\begin{itemize}
    ℹtem \textbf{Focus}: Survival horror + mythic revelation
    ℹtem \textbf{Key Mechanics}: Environmental table, name fragments, Heroine's phantom
    ℹtem \textbf{Pacing}: 3−4 sessions
    ℹtem \textbf{GM Tips}:
        \begin{itemize}
            ℹtem Vary the environmental effects − don't use them every scene
            ℹtem Let the Heroine's phantom shift based on PC actions
            ℹtem Name fragments should feel earned, not random
        \end{itemize}
\end{itemize}

\subsection{Act III: The Fortress Heart}
\begin{itemize}
    ℹtem \textbf{Focus}: Grim revelation + predatory architecture
    ℹtem \textbf{Key Mechanics}: Room generator, adversary table, Heroine's judgment
    ℹtem \textbf{Pacing}: 2−3 sessions
    ℹtem \textbf{GM Tips}:
        \begin{itemize}
            ℹtem The fortress should feel alive and reactive
            ℹtem Use the Heroine as a dynamic antagonist, not static boss
            ℹtem Let the PCs' Memory Track influence her behavior
        \end{itemize}
\end{itemize}

\section{Key Mechanical Reminders}

\subsection{Story Beat Economy}

¢ering
\begin{longtable}{|c|l|l|}
ℎline
\textbf{SB} & \textbf{Direwood Use} & \textbf{Impact} \\
ℎline
1 SB & Minor environmental shift & Path loops, fog thickens \\
2 SB & Moderate complication & Wraith attraction, memory loss \\
3 SB & Serious threat & Shadewalker patrol, trap activation \\
4+ SB & Major escalation & Heroine's wrath, fortress restructuring \\
ℎline
\end{longtable}
∩tion{Direwood Story Beat Applications}


\subsection{Environmental Horror Tags}
\begin{itemize}
    ℹtem \textbf{[MIST−CHOKED]}: Disorientation, impaired perception, fear generation
    ℹtem \textbf{[HAUNTED]}: Reality seams, time loops, memory distortion
    ℹtem \textbf{[STARVED−WORLD]}: Resource scarcity, will−draining, identity erosion
\end{itemize}

\section{Character Archetype Support}

\subsection{Tank/Controller Players}
\begin{itemize}
    ℹtem Emphasize defensive positioning opportunities
    ℹtem Provide clear crowd control scenarios
    ℹtem Reward proactive threat management
\end{itemize}

\subsection{Striker/DPS Players}
\begin{itemize}
    ℹtem Offer tactical advantage positioning
    ℹtem Include mobile combat scenarios
    ℹtem Provide clear damage−dealing opportunities
\end{itemize}

\subsection{Support/Buffer Players}
\begin{itemize}
    ℹtem Create party−cohesion challenges
    ℹtem Include social manipulation opportunities
    ℹtem Provide healing/resource management moments
\end{itemize}

\subsection{Utility/Scout Players}
\begin{itemize}
    ℹtem Design puzzle/problem−solving scenarios
    ℹtem Include exploration−focused challenges
    ℹtem Reward specialized skill usage
\end{itemize}

\section{Thematic Reinforcement}

\subsection{Consistency Tips}
\begin{itemize}
    ℹtem Always tie mechanical effects to narrative causes
    ℹtem Let the forest react to player choices thematically
    ℹtem Use recurring motifs (bones, mist, names, memory)
    ℹtem Maintain the tension between heroism and tragedy
\end{itemize}

\subsection{Horror Pacing}
\begin{enumerate}
    ℹtem \textbf{Establish Normalcy}: Start with recognizable elements
    ℹtem \textbf{Introduce Wrongness}: Small, subtle violations
    ℹtem \textbf{Escalate Gradually}: Build to major violations
    ℹtem \textbf{Provide Relief}: Moments of safety or clarity
    ℹtem \textbf{Return Stronger}: Escalate beyond initial wrongness
\end{enumerate}

\section{Troubleshooting Common Issues}

\subsection{Players Moving Too Slowly}
\begin{itemize}
    ℹtem Increase environmental pressure (Dread Timer ticks)
    ℹtem Introduce time−sensitive threats
    ℹtem Use faction pressure from Payden's Port
\end{itemize}

\subsection{Players Rushing Ahead}
\begin{itemize}
    ℹtem Enforce consequences for missed investigation
    ℹtem Use environmental barriers that require preparation
    ℹtem Let important clues become inaccessible
\end{itemize}

\subsection{Combat Dominance}
\begin{itemize}
    ℹtem Emphasize non−combat solutions to threats
    ℹtem Use adversaries that are difficult to defeat conventionally
    ℹtem Introduce social or psychological consequences
\end{itemize}

\subsection{Investigation Stalling}
\begin{itemize}
    ℹtem Provide multiple paths to information
    ℹtem Use active investigation (visions, dreams, encounters)
    ℹtem Let the environment provide clues organically
\end{itemize}

\section{Optional Variations}

\subsection{Shortened Campaign}
\begin{itemize}
    ℹtem Focus on core mystery (name restoration)
    ℹtem Reduce faction complexity
    ℹtem Streamline environmental challenges
    ℹtem Condense fortress exploration
\end{itemize}

\subsection{Increased Horror Focus}
\begin{itemize}
    ℹtem Emphasize psychological effects
    ℹtem Increase memory distortion elements
    ℹtem Add more personal horror connections
    ℹtem Reduce combat options
\end{itemize}

\subsection{Mythic Heroism Focus}
\begin{itemize}
    ℹtem Emphasize the Heroine's redemption arc
    ℹtem Provide more opportunities for legendary actions
    ℹtem Focus on saga and legacy themes
    ℹtem Reduce horror elements
\end{itemize}

\section{Tracking Sheets and Handouts}

\subsection{Player Tracking}
\begin{itemize}
    ℹtem Memory Track shared sheet
    ℹtem Name Fragment collection log
    ℹtem Faction relationship tracker
    ℹtem Direwood survival log
\end{itemize}

\subsection{GM Tracking}
\begin{itemize}
    ℹtem Dread Timer status
    ℹtem Heroine's attitude tracker
    ℹtem Environmental effect log
    ℹtem Faction pressure indicators
\end{itemize}

\section{Conclusion}
Remember: The Direwood responds to the players' approach. Respect their choices, escalate meaningfully, and let the tragedy of the forgotten heroine drive both horror and hope.

\chapter{Thematic Tier III Character Examples}

\section{Aeler Military Engineer}
\textbf{Name:} Thrain Stonecaller \\
\textbf{Background:} Aeler Charter Apprentice \\
\textbf{Concept:} Fortress−Craft Master who specializes in defensive engineering and siege−breaking \\
\textbf{Attributes:} Body 3, Wits 4, Spirit 3, Presence 2 \\
\textbf{Skills:} Craft 4, Tinker 3, Tactics 3, Lore 2 \\
\textbf{Talents:} Vent Prior's Training (3 XP), Stone−Sense (5 XP), Defensive Architecture (4 XP) \\
\textbf{Patron:} The Sealed Gate \\
\textbf{Familiar:} Thiasos − Ironscale (Stone Sentinel) \\
\textbf{Codex:} Charter of the Deep Vaults \\
\textbf{Assets:} 
\begin{itemize}
    ℹtem Minor: Engineering Kit (surveyor's tools, measuring chains)
    ℹtem Follower: Gunnvor Ironwright (Cap 2 Siege Specialist)
\end{itemize}
\textbf{Background Tags:} Charter−Stamped, Stone−Speech Initiate \\
\textbf{Signature Contact:} Master Builder Emrik (Cap 1) \\
\textbf{Background Boon:} Once/session around structures, ignore the first environmental complication from collapse or strain.

\section{Mistland Local Guide}
\textbf{Name:} Uria Mistweaver \\
\textbf{Background:} Mistlands Bell−Keeper \\
\textbf{Concept:} Bell−tender turned wilderness guide who knows the dangerous paths through the fog \\
\textbf{Attributes:} Body 2, Wits 3, Spirit 4, Presence 3 \\
\textbf{Skills:} Survival 4, Perception 3, Lore 3, Sway 2 \\
\textbf{Talents:} Heightened Senses (3 XP), Mist Reading (4 XP), Bell−Speech (3 XP) \\
\textbf{Patron:} The Silent Choir \\
\textbf{Familiar:} Whisper (Corvid Scout) \\
\textbf{Codex:} Songs of the Dusk Bells \\
\textbf{Assets:}
\begin{itemize}
    ℹtem Minor: Bell−Pouch (ceremonial handbells)
    ℹtem Follower: Bramble Thornwick (Cap 3 Wilderness Guide)
\end{itemize}
\textbf{Background Tags:} Bell−Law, Causeway Rights \\
\textbf{Signature Contact:} Dusk−Warden Morn (Cap 1) \\
\textbf{Background Boon:} Once/session in mist, choose route−or−safety: negate a travel delay or downgrade a hazard.

\section{Zakov Smuggler Captain}
\textbf{Name:} Captain "Shadow" Kestra Voss \\
\textbf{Background:} Underworld Hand \\
\textbf{Concept:} Pirate−turned−smuggler who runs contraband past Aeler checkpoints \\
\textbf{Attributes:} Body 3, Wits 4, Spirit 3, Presence 3 \\
\textbf{Skills:} Sway 4, Ranged 3, Subterfuge 3, Survival 2 \\
\textbf{Talents:} Second Wind (2 XP), Keen Senses (2 XP), Social Chameleon (4 XP), Shipmaster's Instinct (4 XP) \\
\textbf{Patron:} Raéyn (Mistress of the Sea) \\
\textbf{Familiar:} Stormrider (Seabird Scout) \\
\textbf{Codex:} Tide−Rider's Charter \\
\textbf{Assets:}
\begin{itemize}
    ℹtem Minor: Hidden Cargo Compartment
    ℹtem Standard: Smuggling Vessel "Night's Passage"
    ℹtem Follower: Finn the Knife (Cap 3 Deckhand/Scout)
\end{itemize}
\textbf{Background Tags:} Harbor Handsigns, Black Banner Cant \\
\textbf{Signature Contact:} Fence Master Torr (Cap 1) \\
\textbf{Background Boon:} Once/session when using a cant, gain +1d to set up or slip out.

\section{Linnic Trader−Raider}
\textbf{Name:} Bjorn Iceheart \\
\textbf{Background:} Frontier Marcher Homesteader \\
\textbf{Concept:} Seafaring trader who raids when trade winds fail \\
\textbf{Attributes:} Body 4, Wits 3, Spirit 3, Presence 3 \\
\textbf{Skills:} Melee 4, Ranged 3, Survival 3, Command 2 \\
\textbf{Talents:} Berserker Rage (8 XP), Weapon Mastery (4 XP), Cold Weather Survival (3 XP), Naval Tactics (4 XP) \\
\textbf{Patron:} Thrysos (King of Revels) \\
\textbf{Familiar:} Frostfang (Wolf Companion) \\
\textbf{Codex:} Saga of the Whale−Road \\
\textbf{Assets:}
\begin{itemize}
    ℹtem Minor: Dragon Boat "Stormbreaker"
    ℹtem Standard: Trading Post in Linnstad
    ℹtem Follower: Sigrun Ironscale (Cap 3 Warrior−Trader)
\end{itemize}
\textbf{Background Tags:} Veteran−of−the−Marches, Muster Papers \\
\textbf{Signature Contact:} Clan−Chief Harald (Cap 1) \\
\textbf{Background Boon:} Once/session in organized violence, ignore the first point of situational Disadvantage from confusion or fear.

\section{Lethai−al Valewood Ranger}
\textbf{Name:} Aelindra Greenwalker \\
\textbf{Background:} Rural Valewood Forager \\
\textbf{Concept:} Forest guardian who protects ancient paths from outsiders \\
\textbf{Attributes:} Body 3, Wits 4, Spirit 3, Presence 2 \\
\textbf{Skills:} Nature 4, Stealth 3, Perception 3, Lore 3 \\
\textbf{Talents:} Canopy Spring (4 XP), Scent of Rain (4 XP), Green Speech (3 XP), Arrow−Song (4 XP) \\
\textbf{Patron:} Grimmir (Wild Speaker) \\
\textbf{Familiar:} Thornwhisper (Forest Cat) \\
\textbf{Codex:} Songs of Root and Branch \\
\textbf{Assets:}
\begin{itemize}
    ℹtem Minor: Hunter's Bow "Whisperleaf"
    ℹtem Follower: Moss (Cap 2 Forest Guide/Healer)
\end{itemize}
\textbf{Background Tags:} Green Courts Token, River−Ferry Nod \\
\textbf{Signature Contact:} Leaf−Warden Sylvan (Cap 1) \\
\textbf{Background Boon:} Once/session in the wild, downgrade a Travel complication by one step.

\section{Aelinnel Gnome Artificer}
\textbf{Name:} Quillwick Coppermeasure \\
\textbf{Background:} Academic of the Dialogues \\
\textbf{Concept:} Meticulous inventor who creates magical devices using precise calculations \\
\textbf{Attributes:} Body 2, Wits 5, Spirit 2, Presence 3 \\
\textbf{Skills:} Craft 4, Arcana 3, Investigation 3, Lore 3 \\
\textbf{Talents:} Long Memory (3 XP), Number Music (4 XP), Clockwork Mastery (6 XP), Two−Ledger Talk (3 XP) \\
\textbf{Patron:} Sacred Geometry \\
\textbf{Familiar:} Cogwheel (Mechanical Companion) \\
\textbf{Codex:} Theorems of Perfect Proportion \\
\textbf{Assets:}
\begin{itemize}
    ℹtem Minor: Workshop with Precision Tools
    ℹtem Standard: Calculating Engine (Minor Artifact)
    ℹtem Follower: Tink (Cap 2 Apprentice/Assistant)
\end{itemize}
\textbf{Background Tags:} School−Sealed, Reading Privileges \\
\textbf{Signature Contact:} Lecturer Brightmind (Cap 1) \\
\textbf{Background Boon:} Once/session in research or disputation, ask one clarifying question the GM must answer plainly.

\section{Aelaerem Halfling Herbalist}
\textbf{Name:} Bramble Goodwort \\
\textbf{Background:} Rural Valewood Forager \\
\textbf{Concept:} Gentle healer who knows the dangerous plants of borderlands \\
\textbf{Attributes:} Body 2, Wits 3, Spirit 4, Presence 3 \\
\textbf{Skills:} Medicine 4, Nature 3, Perception 3, Sway 2 \\
\textbf{Talents:} Heightened Senses (3 XP), Root−Balance (3 XP), Healing Touch (4 XP), Herb Lore (3 XP) \\
\textbf{Patron:} The Silent Choir \\
\textbf{Familiar:} Sage (Garden Snake) \\
\textbf{Codex:} Remedies of the Hearth \\
\textbf{Assets:}
\begin{itemize}
    ℹtem Minor: Herb Garden and Apothecary Kit
    ℹtem Follower: Pip Willowfoot (Cap 2 Gatherer/Apprentice)
\end{itemize}
\textbf{Background Tags:} Neighborhood Trust, Guild−Friend \\
\textbf{Signature Contact:} Village Elder Rose (Cap 1) \\
\textbf{Background Boon:} Once/session among familiar folk, convert a failure on routine labor to a success−with−cost.


\end{document}
