\documentclass[12pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{multicol}

% Define colors
\definecolor{silkpurple}{RGB}{128, 0, 128}
\definecolor{silkgold}{RGB}{218, 165, 32}
\definecolor{silkred}{RGB}{178, 34, 34}

% Section formatting
\titleformat{\section}
{\color{silkpurple}\normalfont\Large\bfseries}
{\color{silkpurple}\thesection}{1em}{}

\titleformat{\subsection}
{\color{silkgold}\normalfont\large\bfseries}
{\color{silkgold}\thesubsection}{1em}{}

% Item spacing
\setlist{itemsep=2pt}

\title{\color{silkpurple} \textbf{The Merchant War}\\ \large An Adventure for Fate's Edge - Silkstrand}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Adventure Overview}

\textbf{Title:} The Merchant War\\
\textbf{Region:} Silkstrand - City of Bridges \& Dyewater\\
\textbf{Theme:} Economic warfare, social manipulation, urban intrigue\\
\textbf{Level:} Seasoned (Tier II)\\
\textbf{Length:} 3-4 Sessions

\section{Premise}

Two powerful merchant houses in Silkstrand are engaged in an escalating economic war that's disrupting trade throughout the region. What began as competitive business practices has evolved into sabotage, assassination, and manipulation of the city's political structure. The players must navigate this conflict while pursuing their own objectives, choosing sides or finding a way to end the war before it destroys the city's economy entirely.

\section{Hook}

The adventure begins when the players encounter one of the following scenarios:

\begin{itemize}
  \item Their cargo is seized at the docks due to a "regulatory dispute" between the two merchant houses, requiring them to navigate the political maze to recover their goods.
  \item A contact offers them a lucrative job that involves retrieving sensitive information from the rival merchant house, but the job goes wrong and they become targets.
  \item They witness an assassination attempt on a prominent merchant, making them both witnesses and potential targets in the escalating conflict.
\end{itemize}

\section{Key Factions}

\subsection{House Valdris}
Led by Guildmistress Thalia Valdris, this traditional merchant house controls the dye trade and has strong connections with the old nobility. They favor established procedures and view the upstart Crimson Sails as disrespectful newcomers threatening the established order.

\textbf{Strengths:} Established connections, dye monopoly, political influence\\
\textbf{Methods:} Legal maneuvering, political pressure, traditional sabotage\\
\textbf{Symbol:} Purple and silver

\subsection{Crimson Sails Trading Company}
Led by Captain-Merchant Renn Kestral, this aggressive newcomer has disrupted traditional trade routes with innovative practices and ruthless efficiency. They view House Valdris as corrupt and stagnant, blocking progress for personal gain.

\textbf{Strengths:} Naval power, innovative practices, popular with younger merchants\\
\textbf{Methods:} Economic warfare, bribery, direct action\\
\textbf{Symbol:} Crimson and gold

\section{Key NPCs}

\subsection{Guildmistress Thalia Valdris}
A calculating woman in her fifties who has built House Valdris into a Silkstrand institution. She believes in tradition and proper procedure, but her methods have become increasingly desperate as Crimson Sails gains ground. She offers the players legitimate contracts with the promise of long-term stability.

\subsection{Captain-Merchant Renn Kestral}
A charismatic former corsair who turned legitimate trader. In his forties, he's dynamic and innovative, willing to take risks that traditional merchants wouldn't consider. He offers the players exciting opportunities with high rewards but significant danger.

\subsection{Matron Esme Blackwater}
The nominal ruler of Silkstrand who maintains order while the merchant houses fight their war. She's growing increasingly concerned about the impact on the city's stability and may approach the players to help end the conflict.

\subsection{The Broker}
A mysterious figure who facilitates transactions between the two houses while secretly fueling the conflict for profit. The Broker knows secrets about both factions and can provide valuable information to the players - for a price.

\section{Key Locations}

\subsection{Three-Queens Bridge}
The central marketplace where the two factions compete for customers. The bridge has become a battleground with each house trying to outmaneuver the other through pricing, quality, and customer service. Duels are common here when business disputes escalate.

\subsection{The Dye Works}
House Valdris's primary facility where they produce their famous silk dyes. The works are heavily guarded and contain valuable trade secrets. Recent sabotage attempts have made security even tighter.

\subsection{The Crimson Docks}
Crimson Sails' naval base where their fast ships are maintained. The docks are a hub of activity and a potential target for House Valdris operations. The area is heavily patrolled and has its own code of conduct.

\subsection{The Archivolt}
The bureaucratic heart of Silkstrand where contracts are recorded and disputes are settled. Both merchant houses maintain offices here and employ advocates to argue their cases. The building is a maze of legal precedents and bureaucratic procedures.

\subsection{The Salt Gate Quay}
The main import/export facility where goods enter and leave Silkstrand. Control of the quay means control of trade flows, making it a key strategic location. Recent "accidents" have disrupted operations and increased costs for both houses.

\section{Plot Structure}

\subsection{Session 1: Entanglement}
The players become involved in the merchant war through their hook. They should:
\begin{itemize}
  \item Experience the effects of the economic warfare firsthand
  \item Meet representatives from both factions
  \item Learn about the conflict's origins and current state
  \item Make initial choices about which side to support (if any)
\end{itemize}

\subsection{Session 2: Deepening Involvement}
The players take on more significant roles in the conflict. They should:
\begin{itemize}
  \item Complete missions for their chosen faction (or both)
  \item Experience the escalating consequences of the war
  \item Discover secrets about both houses
  \item Face moral dilemmas about their involvement
\end{itemize}

\subsection{Session 3: The Breaking Point}
The conflict reaches a crisis point that threatens to destroy both factions and the city. The players should:
\begin{itemize}
  \item Deal with the full consequences of their previous actions
  \item Make crucial decisions about how to resolve the conflict
  \item Potentially face off against the secret manipulator
  \item Determine the future of Silkstrand's merchant community
\end{itemize}

\section{Key Mechanics}

\subsection{Merchant War Clock [8]}
Tracks the escalating conflict between the two merchant houses. Advances when:
\begin{itemize}
  \item Economic sabotage is successful
  \item Assassinations occur
  \item Political maneuvering fails
  \item Players make choices that escalate the conflict
\end{itemize}

\subsection{Reputation Tracks}
Both merchant houses track their opinion of the players:
\begin{itemize}
  \item House Valdris: Traditionalist to Rebel
  \item Crimson Sails: Corrupt to Progressive
\end{itemize}
Player actions shift these tracks, affecting available missions and NPC reactions.

\subsection{Bridge Duel Protocol}
When business disputes escalate to violence on Three-Queens Bridge:
\begin{itemize}
  \item Declare Stakes: Purse, Permit, or Pride
  \item Seconds may spend 1 Boon to Call the Colors
  \item Refusing triggers automatic reputation damage
\end{itemize}

\subsection{Exchange Mini-Game}
Trading and negotiation mechanics:
\begin{itemize}
  \item Track a scene-long Spread (starts at 1)
  \item Each good call on the floor moves Spread by ±1
  \item Cash Out: gain Price Edge tag (+1 die) when buying/selling
\end{itemize}

\section{Possible Resolutions}

\subsection{Faction Victory}
The players help one faction decisively defeat the other:
\begin{itemize}
  \item Complete a major sabotage mission against the rival
  \item Win a crucial legal case in the Archivolt
  \item Eliminate key leadership of the opposing faction
\end{itemize}
\textbf{Consequences:} Winner gains control, but the city's balance of power shifts dangerously.

\subsection{Negotiated Settlement}
The players broker a peace agreement between the factions:
\begin{itemize}
  \item Discover the Broker's manipulation and expose them
  \item Find a compromise that benefits both houses
  \item Convince the Matron to intervene with authority
\end{itemize}
\textbf{Consequences:} Stable business environment, but both houses owe the players favors.

\subsection{Revolutionary Change}
The players help overthrow the old system entirely:
\begin{itemize}
  \item Expose corruption in both houses
  \item Rally the common merchants and workers
  \item Install a new form of governance
\end{itemize}
\textbf{Consequences:} Complete upheaval with uncertain future, but potential for great personal power.

\section{Rewards and Consequences}

\subsection{Immediate Rewards}
\begin{itemize}
  \item Bridge token for peak-hour cart crossing
  \item Dye-permit chit to process "questionable" color
  \item Warehouse seal from winning faction
  \item Exchange pass for trading without sponsor
  \item Watergate priority to skip closures
\end{itemize}

\subsection{Long-term Consequences}
\begin{itemize}
  \item Permanent business relationships with the surviving faction
  \item Reputation as either a stabilizing force or dangerous revolutionary
  \item Access to exclusive trading opportunities or loss of previous connections
  \item Potential leadership roles in the new order or exile from Silkstrand
\end{itemize}

\section{GM Notes}

\subsection{Social Dynamics}
Emphasize the social aspects of the conflict. This is as much about reputation and influence as it is about money. Every interaction should have social stakes and potential long-term consequences.

\subsection{Economic Realities}
Make the economic warfare feel real by showing its effects on ordinary people. Rising prices, job losses, and disrupted supply chains should be visible throughout the city.

\subsection{Moral Ambiguity}
Neither faction is entirely good or evil. House Valdris provides stability but resists progress, while Crimson Sails brings innovation but disrupts established communities. Players should face genuine moral choices.

\subsection{Player Agency}
Provide multiple paths to influence the conflict. Players might choose to work for one side, play both against each other, or work to end the war entirely. Their choices should have meaningful consequences.

\subsection{Urban Environment}
Use Silkstrand's unique environment - the bridges, dyeworks, and waterways - to create distinctive scenes. The city itself should feel like a character in the story.

\subsection{Story Beats}
Use the economic conflict to generate Story Beats that complicate the players' efforts while advancing the plot. Sabotage, betrayal, and unexpected alliances should keep the players on their toes.

\end{document}
