\documentclass[18pt, letterpaper, twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{titling}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{hyperref} % Should be loaded last
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={The Nameless},
    pdfauthor={Fate's Edge Community},
    pdfsubject={A Fate's Edge Adventure},
    pdfcreator={LaTeX with hyperref},
    pdfproducer={pdflatex}
}

% Page Layout
\geometry{
    a4paper,
    left=25mm,
    right=25mm,
    top=25mm,
    bottom=30mm
}

% Header & Footer
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[LO]{The Nameless}
\fancyhead[RE]{Fate's Edge Adventure}
\renewcommand{\headrulewidth}{0.4pt}
\fancyfoot[C]{}

% Title Format
\titleformat{\section}{\Large\bfseries\filcenter}{}{0em}{}
\titleformat{\subsection}{\large\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\bfseries}{}{0em}{}

% Paragraph Spacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5em}

% Custom Environments
\newenvironment{floorbox}[1]{%
    \begin{center}
    \begin{minipage}{0.95\textwidth}
    \subsection*{#1}
    \vspace{-0.5em}
    \rule{\textwidth}{0.5pt}
    \vspace{0.5em}
}{%
    \vspace{0.5em}
    \rule{\textwidth}{0.5pt}
    \end{minipage}
    \end{center}
}

\newenvironment{gmnote}[1]{%
    \vspace{0.5em}
    \noindent\textbf{#1}
    \begin{quote}
}{%
    \end{quote}
    \vspace{0.5em}
}

\newenvironment{playerbox}[1]{%
    \vspace{0.5em}
    \noindent\textbf{#1}
    \begin{quote}
}{%
    \end{quote}
    \vspace{0.5em}
}

\newenvironment{resolutionbox}{%
    \vspace{0.5em}
    \noindent\textbf{Resolution Fork:}
    \begin{quote}
}{%
    \end{quote}
    \vspace{0.5em}
}

\newenvironment{boxedtext}[1]{%
    \vspace{0.5em}
    \noindent\textbf{#1}
    \begin{quote}
}{%
    \end{quote}
    \vspace{0.5em}
}

\newcommand{\memoryline}[1]{\noindent\textbf{Memory in One Line:} #1\par}
\newcommand{\tonetags}[1]{\noindent\textbf{Tone Tags:} #1\par}
\newcommand{\startposition}[1]{\noindent\textbf{Start Position:} #1\par}
\newcommand{\floorclock}[1]{\noindent\textbf{Floor Clock:} \fbox{#1}\par}
\newcommand{\handles}[1]{\noindent\textbf{Handles:} #1\par}
\newcommand{\sbitem}[2]{\item[#1 SB:] #2}
\newcommand{\deckdraw}[1]{\noindent\textbf{Deck of Consequences:} #1\par}
\newcommand{\keyname}[1]{\noindent\textbf{Key Name:} #1\par}
\newcommand{\nameimportance}[1]{\noindent\textbf{Why This Name Matters:} #1\par}
\newcommand{\boon}[1]{\noindent\textbf{Boon:} #1\par}
\newcommand{\burden}[1]{\noindent\textbf{Burden:} #1\par}
\newcommand{\entrytext}[1]{\begin{boxedtext}{Entry Read-Aloud:} #1 \end{boxedtext}}
\newcommand{\exittext}[1]{\begin{boxedtext}{Exit Beat:} #1 \end{boxedtext}}

\newcommand{\patrontitle}[1]{\section*{#1}}
\newcommand{\tenet}[2]{\noindent\textbf{#1.} #2\par}
\newcommand{\devotion}[2]{\noindent\textbf{Favor #1 --- #2}\par}
\newcommand{\rite}[2]{\noindent\textbf{#1:} #2\par}
\newcommand{\compel}[1]{\noindent\textbf{Compel:} #1\par}
\newcommand{\taboo}[1]{\noindent\textbf{Taboo:} #1\par}
\newcommand{\offering}[1]{\noindent\textbf{Offering:} #1\par}
\newcommand{\obligation}[1]{\noindent\textbf{Obligation:} #1\par}
\newcommand{\synergy}[1]{\noindent\textbf{Synergy:} #1\par}

\newcommand{\corruptiontitle}[1]{\subsection*{#1}}
\newcommand{\threshold}[2]{\noindent\textbf{#1+ Corruption:} #2\par}

\begin{document}

% Title Page
\begin{center}
    {\Huge \textbf{The Nameless}}\\[1em]
    {\Large A tragic, choice-driven dungeon crawl for Fate's Edge}\\[2em]
    \rule{\textwidth}{1pt}\\[1em]
    {\large \textit{Version 1.0 - Compiled from Penance.txt}}\\[0.5em]
    {\large \textit{System: Fate's Edge SRD}}\\[1em]
    \rule{\textwidth}{1pt}
\end{center}

\thispagestyle{empty}
\newpage
\setcounter{page}{1}

% Table of Contents
\tableofcontents
\newpage

% How to Run This Adventure
\section{How to Run This Adventure}

\subsection{Floor-as-Leg Procedure}
\begin{enumerate}[leftmargin=*]
    \item \textbf{Seed the Floor:} Treat each floor as a Travel ``leg.'' Draw your seed (Spade = place, Heart = actor, Club = pressure, Diamond = leverage) from a reskinned ``Dungeon Deck,'' or choose elements that match the listed floor. Highest rank sets the Floor Clock.
    \item \textbf{Set Position:} Start at Dominant. Adjust during play for bold risks, insight, or rising backlash.
    \item \textbf{Track the Floor Clock [4--6]:} Use [4] for compact scenes, [5] for tense puzzles/social play, [6] for complex combats or multifaceted dilemmas.
    \item \textbf{SB Economy = Revelation Engine:} Prefer spending SB to reveal memory, raise stakes, or tilt loyalties over raw damage. Draw from the Deck of Consequences when the scene earns it (multiple 1s, spikes in tension) to color omens, debts, or emotional fallout.
    \item \textbf{End with a Choice:} Each floor ends on a fork that reframes the lich. Choices tune Position for the next floor and unlock boons or burdens.
\end{enumerate}

\subsection{Deck Usage in the Dungeon}
\begin{description}[leftmargin=*, labelwidth=3cm, style=standard]
    \item[Hearts:] Shame, love, vows, memory-ghosts.
    \item[Clubs:] Breakage, traps, relentless pursuers.
    \item[Diamonds:] Wards, keys, rights of passage, bargains.
    \item[Spades:] Necromancy, omens, soul-anchors, whispers.
\end{description}

\subsection{XP \& Boons}
\begin{itemize}[leftmargin=*]
    \item Award XP for discoveries, hard choices, embracing costs, and major objectives.
    \item Grant Boons for partials/misses and compassionate play; allow Boons to steady Position or reroll at pivotal moments.
\end{itemize}

\subsection{Final Persuasion Clock}
The last floor is resolved via a Persuasion Clock [6--8] with three destinations: Continue Penance, Find Peace, Join the Living. Table and player choices across prior floors shift starting Position and DV.

\subsection{Tone \& Safety}
This adventure centralizes grief, broken vows, and contrition. Check lines/veils, use safety tools, and invite players to define what reconciliation looks like in this world.

\subsection{Tiering \& Difficulty Dials}
\begin{description}[leftmargin=*, labelwidth=4cm, style=standard]
    \item[Party Tier I--II:] Default clocks [4--5]; sparse reinforcements; SB favors revelations and single-issue conditions.
    \item[Tier III+:] More [6] clocks, layered wards, and factional crossfire; SB can call reinforcements, gear failure, or scene shifts in addition to revelations.
\end{description}

% Entry Hook
\section{Entry: The Forbidden Library (Hook)}
A silent archive under municipal seal; brittle chains anchor a violet door. Research exposes this dungeon's purpose: a penitential labyrinth. Breaking the seal opens to the Landing Stair and a breeze like the last breath of a vow.

\textbf{Skill Hooks:} Lore (necromantic jurisprudence), Insight (contrition vs. cruelty), Tinker (seal-hardware), Arcana (sympathy lines).

\textbf{Stakes:} You descend to hold the lich accountable---or to free him.

% The 30 Floors of Penance
\section{The 30 Floors of Penance}

Each floor below lists: Memory, Environment, Clock, Handles, SB Spends, and Choices. Start Position is Dominant unless stated.

% --- Floor 1 ---
\begin{floorbox}{1 --- The Abandoned Apprentice}
\memoryline{He left his first student locked out of the rite to claim credit.}
\tonetags{contrition • workshop • betrayal • stolen knowledge}
\startposition{Dominant}
\floorclock{[5] --- Tools hum louder; chalk-circle sparks; memory-constructs awaken.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} He believed mastery could only be earned through trial by fire, not gentle guidance. The student was weak, he told himself---better to let them fail and learn resilience than to coddle them into dependency. But the truth was simpler: he feared being surpassed, so he denied his apprentice access to the ritual that would have made them both powerful. The chalk-circle still bears scorch marks from where the rejected initiate tried to breach it. During the Siege of Ecktoria, such petty jealousies would prove costly when collaboration was needed most.

\textbf{From the Harmed Party's Perspective:} The apprentice had waited years for this moment---years of servitude, of fetching components and cleaning vessels, all for a chance to finally become something more than a glorified servant. Denied entry, they fled into the night, their dreams of greatness reduced to ash. What justice would have looked like then was simple: inclusion, recognition, a fair chance to prove themselves. In the years that followed, many apprentices would flee the collapsing academies as the empire fell.

\textbf{Penance Sought:} To acknowledge his apprentice's worth and give them the credit they were denied. The lich wonders if his refusal to share knowledge contributed to the empire's intellectual decline.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} A workshop bisected by a glowing chalk-circle; arcane implements hum with residual charge. Ancient textbooks bear the cracked laurel-circle seals of the Utaran academies.

\textbf{Anchors:} Chalk-circle, apprentice's notes, master's grimoire bearing Utaran administrative markings, construct servitors

\textbf{Senses:} The scent of ozone and burnt parchment; the faint sound of weeping echoing from within the circle; a distant echo of the Noon Bell that once marked time in the empire's great halls.

\textbf{Adversaries/Agents:} Memory-constructs of failed apprentices, spectral echoes of rejected students, the ghost of an Utaran administrator demanding accountability.
\end{gmnote}

\handles{Arcana (counter the glyph's binding pattern, DV 3) • Sway (validate the apprentice-echo's grievances, DV 2) • Tinker (ground the circle's energy discharge, DV 3) • Skirmish (shatter the memory-constructs, DV 2)}

\textbf{Position Drift:} Mercy (validating the apprentice) raises Position next floor; cruelty (destroying evidence) drops it.

\textbf{When the Floor Clock Fills:} The circle erupts in wild energy, summoning additional constructs while the apprentice's notes begin to burn themselves unreadable. A fragment of Utaran law scrolls out, bearing regulations on proper master-apprentice relations that were ignored during the empire's final years.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{The apprentice's notes accuse a small theft of credit. A marginal note references "the old ways before the siege."}
    \sbitem{2}{Ward backlash binds a PC's shadow to the circle. The shadow whispers of bureaucratic failures that preceded the empire's fall.}
    \sbitem{3}{A hidden chamber reveals the true extent of the apprentice's potential. Among their designs are improved fortification plans that might have helped during the Siege of Ecktoria.}
    \sbitem{4+}{The master's grimoire opens to a page showing the ritual the apprentice was denied. A bookmarked page contains a letter from the Utaran Ministry of Defense requesting such rituals for wartime applications.}
\end{description}

\deckdraw{Draw when the circle's energy becomes unstable or when a PC's shadow is bound---omens of denied potential, debts of unfulfilled promises, echoes of empire's collapse.}

\begin{gmnote}{The Name-as-Key}
Each floor is locked by a true name tied to the betrayal. Discovering and correctly using it is the cleanest exit. Combat or workarounds are possible but risk harsher downstream costs.

\keyname{ELARA THORNWRIGHT, FIRST STUDENT}
\nameimportance{Restoring her full name acknowledges she was more than the wound she became. Her family name connects to Utaran nobility who fled the empire's collapse.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of spell components spell her name, with a component list dated to the years before the Siege of Ecktoria
    \item Palimpsest: Her name visible only under the light of the circle's energy, revealed beneath erased Utaran military contracts
    \item Emotional Key: Appears when a PC chooses to validate her achievements over the master's reputation, echoing the Utaran values of merit over birth
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Spell components arranged in her initials, with a component tag noting its origin from Utaran supply caches
    \item[Actionable Lead:] Arcana reveals energy traces leading to her hidden potential, traces that match patterns used in Utaran defensive magic
    \item[Aha! Proof:] Tinker work uncovers her secret laboratory with superior designs, including annotations comparing her work to official Utaran military specifications
\end{description}

\textbf{Extraction Test:} Evidence or Arcana to trace her contributions (on fail, the circle's energy intensifies but reveals her middle name and a connection to Utaran refugee networks).

\textbf{Verification \& Use:} Speak her name while placing her notes within the circle; mispronunciations crack the circle (harm 1) and add a clock segment. Speaking her full name correctly causes the circle to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Validate the apprentice's worth (boon: Student's Favor - once: steady Position when learning); Position +1 next floor

\textbf{Restitution:} Return her notes and acknowledge her contributions publicly (Debt 1 - owed to her surviving family, who became Utaran refugees)

\textbf{Renunciation:} Break the circle and erase all records (lose time, gain Stain - Unworthy Teacher)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Student's Favor (once: steady Position when learning)}
\burden{Stain - Unworthy Teacher if you erase records; Ledger - Helped: Elara if you validate her}
\textbf{Shard Kit:} Apprentice's notes grant +1d to sincere apologies later; the notes bear Utaran administrative stamps
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her family's refugee story connects to Utaran's fall
\textbf{Name Ledger Update:} Harmed: Elara Thornwright • Helped: [ ] • Unresolved: [ ]

\entrytext{The air crackles with stolen power as a chalk-circle splits the workshop in two. On one side, pristine tools await an absent master; on the other, scorch marks and scattered notes speak of a student who was never meant to return. Among the debris, fragments of Utaran law bear witness to forgotten obligations.}

\exittext{The circle's light dims as Elara's name echoes through the chamber. Her notes flutter into a neat stack, and for a moment, you feel the weight of all the knowledge that was never shared. The distant echo of the Noon Bell fades with satisfaction.}
\end{floorbox}

% --- Floor 2 ---
\begin{floorbox}{2 --- The Broken Vow}
\memoryline{He abandoned his wedding mid-ritual for discovery.}
\tonetags{contrition • frost • ceremony • whispers}
\startposition{Dominant}
\floorclock{[6] --- Shards of cold fall; vows unravel; guests awaken hostile.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} He believed a perfect theorem could spare future grief, and so he traded the present for the promise. She stood alone with a vow in her mouth and a ring gone cold. The aisle's frost is the breath of every apology he never learned to make. In his pursuit of forbidden knowledge during the empire's final years, he abandoned even his most personal commitments.

\textbf{From the Harmed Party's Perspective:} She had waited years for this day, weaving her hopes into every stitch of her dress. When he vanished mid-ceremony, she was left holding a broken promise and a ring that would never warm. Justice would have been completion---his presence, his commitment, his choice to honor what he had sworn. Many such abandoned ceremonies marked the empire's decline as officials vanished into war efforts.

\textbf{Penance Sought:} To witness a vow completed without him---and bless it. The lich recognizes that his personal failures mirrored the empire's abandonment of its people.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} A chapel of ice, pews under hoarfrost; a choir that sings syllables backward; a ring on a pedestal sweating cold. The architecture bears the hallmarks of late Utaran construction---triple-course river-stone masonry and laurel-circle motifs.

\textbf{Anchors:} Frost-covered altar bearing Utaran civic seals, wedding ring, inverted vows, backward-singing choir, fragments of Utaran marriage law

\textbf{Senses:} The scent of winter roses; the sound of vows being unsaid; cold that seeps into bone; the distant echo of the Noon Bell that once sanctified Utaran ceremonies.

\textbf{Adversaries/Agents:} Memory-echo bride, frost wights, inverted wedding guests, the ghost of a Utaran marriage registrar demanding proper procedure.
\end{gmnote}

\handles{Presence (ask consent to continue the rite, DV 2) • Rites (bind a surrogate witnessing, DV 3) • Wits (trace the escape sigil, DV 2) • Duel (hold off jealous guardians, DV 3)}

\textbf{Position Drift:} Mercy (honoring the vow) raises Position next floor; rejection (breaking the ritual) drops it.

\textbf{When the Floor Clock Fills:} The chapel fractures into cold shards; guests awaken hostile, demanding completion of what was started. A shard of ice reveals a Utaran edict suspending marriage ceremonies during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Petals turn to cutting glass. Among the debris, a Utaran marriage certificate dated just before the Siege of Ecktoria.}
    \sbitem{2}{Choir turns accusatory, singing fragments of Utaran civic hymns about duty to family and state.}
    \sbitem{3}{The ring binds a PC to a test of fidelity. The binding magic resonates with Utaran oath-taking rituals.}
    \sbitem{4+}{A spectral groom arrives to claim his bride. He wears the uniform of a Utaran officer who died in the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the ring binds a PC or when guests turn hostile---omens of broken promises, debts of unfulfilled commitments, echoes of Utaran civic collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{ALYSSA WHITE-LILY}
\nameimportance{Restoring her full name acknowledges she was more than the wound she became. Her family name connects to Utaran nobility who lost everything in the empire's fall.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of guest ledgers spell her name, with ledgers bearing Utaran administrative numbering systems
    \item Backmask: Spoken backward by echo-choir; correct with counter-chant, the choir singing in the style of Utaran ceremonial choirs
    \item Taboo Hearing: Name audible only after an apology / act of restitution, revealed when a PC speaks words of commitment as Utaran citizens once did
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Lilies carved on the aisle, arranged in the pattern of Utaran noble family crests
    \item[Actionable Lead:] Choir sings a phrase that becomes legible when the ring warms, the phrase being a fragment of Utaran marriage vows
    \item[Aha! Proof:] Steam reveals the palimpsest on the lectern, showing her name written over erased Utaran military dispatches
\end{description}

\textbf{Extraction Test:} Evidence or Arcana to restore the lectern's palimpsest (on fail, the frost spreads but reveals her middle name and a connection to Utaran refugee networks).

\textbf{Verification \& Use:} Speak her name while placing the ring on the pedestal's warmer; mispronunciations crack the ring (harm 1) and add a clock segment. Speaking her full name correctly causes the chapel to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Bless a new vow (boon: Blessing vs undead); Position +1

\textbf{Restitution:} Donate a treasured item to her family row (Debt 1, her family having become Utaran refugees)

\textbf{Renunciation:} Break the escape sigil and stay through the ritual (lose time, gain Public Grace later)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Bride's Favor (once: steady Position)}
\burden{Cold Curse if you reject; Ledger - Helped: Alyssa}
\textbf{Shard Kit:} Frosted ring grants +1d to sincere apologies later; the ring bears Utaran noble markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her family's refugee story connects to Utaran's fall
\textbf{Name Ledger Update:} Harmed: Alyssa White-Lily • Helped: [ ] • Unresolved: [ ]

\entrytext{Frost-laced petals hang in the air, refusing to fall. The choir inhales a vow backward, their voices carrying the cadence of Utaran ceremonial traditions now lost to time.}

\exittext{The frost melts into dew; lilies breathe; the aisle warms under your steps. The distant echo of the Noon Bell fades with satisfaction, and for a moment, you hear the sound of proper Utaran ceremonies being conducted with care.}
\end{floorbox}

% --- Floor 3 ---
\begin{floorbox}{3 --- The Fallen Comrade}
\memoryline{He fled, leaving a shield-mate to die.}
\tonetags{contrition • sand • battle • guilt}
\startposition{Dominant}
\floorclock{[5] --- Banners tighten into nooses; sand shifts to quicksand; campaign clock ticks.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The battle was lost---their position compromised, surrounded by enemies who showed no mercy. He told himself it was tactical retreat, strategic withdrawal, but deep down he knew the truth: fear had driven him to abandon his comrade. The sand still holds the weight of that choice, and the banners remember every fallen ally he left behind. During the Siege of Ecktoria, such cowardice would prove epidemic as Utaran forces faced impossible odds.

\textbf{From the Harmed Party's Perspective:} They had fought side by side for years, sharing rations, covering each other's backs, trusting with their lives. When the moment of greatest need came, their shield-brother vanished, leaving them to face certain death alone. Justice would have been standing together---or dying together. Many Utaran soldiers faced similar betrayals as their empire crumbled.

\textbf{Penance Sought:} To acknowledge the debt owed to his fallen comrade and carry their memory forward. The lich wonders if his abandonment contributed to the collapse of military discipline during Utaran's final campaigns.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Sand-choked battlefield; banners as nooses; scattered weapons and armor half-buried. The landscape bears the scars of ancient warfare, with river-stone fortifications crumbling in the distance.

\textbf{Anchors:} Fallen comrade's shield bearing Utaran military markings, campaign standard with laurel-circle crest, scattered weapons, memorial cairn, fragments of Utaran military dispatches

\textbf{Senses:} The taste of dust and blood; the sound of distant battle cries; shifting sand that threatens to bury everything; the distant echo of Utaran military bugles that once called soldiers to formation.

\textbf{Adversaries/Agents:} Sand wights, memory-bandits, spectral echoes of fallen soldiers, the ghost of a Utaran military tribunal demanding accountability.
\end{gmnote}

\handles{Command (rally ghost cohort, DV 3) • Medicine (salvage the dying echo, DV 2) • Skirmish (cut through memory-bandits, DV 2) • Move (dodge sand-wights, DV 3)}

\textbf{Position Drift:} Staying to hold the line raises Position next floor; abandoning the post drops it.

\textbf{When the Floor Clock Fills:} The battlefield transforms into quicksand; banners tighten into nooses; the old campaign clock begins ticking again. A sand-covered Utaran military standard emerges, bearing the seal of the Ecktoria garrison.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Confusion fog rolls in. Within the fog, ghostly figures of Utaran soldiers who died in the Siege of Ecktoria can be glimpsed.}
    \sbitem{2}{Old campaign clock ticks---reinforcements arrive. The reinforcements wear the uniforms of Utaran auxiliary forces that were sacrificed during the empire's final campaigns.}
    \sbitem{3}{A dying echo whispers the location of hidden supplies. Among the supplies are Utaran military rations and propaganda leaflets about the glory of defending the empire.}
    \sbitem{4+}{The fallen comrade's spirit rises, demanding explanation. The spirit wears the rank insignia of a Utaran officer who died in the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the battlefield becomes quicksand or when the fallen comrade's spirit rises---omens of battlefield guilt, debts of unfulfilled duty, echoes of Utaran military collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THANE IRONHEART}
\nameimportance{Honoring his name acknowledges the sacrifice he made and the debt that remains unpaid. His family name connects to Utaran military tradition, and his service record shows he was among those who died defending Ecktoria.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Syllable Scavenger: Fragments of his name found on scattered weapons and armor, with military tags bearing Utaran serial numbers
    \item Palimpsest: His name visible only when blood is spilled on his shield, revealed beneath erased Utaran commendation citations
    \item Emotional Key: Appears when a PC chooses loyalty over self-preservation, echoing the Utaran military values of duty and honor
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] His initials carved into his shield, arranged in the pattern of Utaran military unit designations
    \item[Actionable Lead:] Medicine work reveals his last words scratched into the sand, words that reference holding the line during the Siege of Ecktoria
    \item[Aha! Proof:] Command ability rallies spirits who speak his name, including other Utaran soldiers who died in the same campaign
\end{description}

\textbf{Extraction Test:} Command or Medicine to recover his final message (on fail, sand buries more evidence but reveals his middle name and military rank).

\textbf{Verification \& Use:} Speak his name while placing his shield on the memorial cairn; mispronunciations cause the sand to shift violently (harm 1) and add a clock segment. Speaking his full name correctly causes the battlefield to resonate with the tone of Utaran military bugles.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Stay and hold the line (Position +1)

\textbf{Restitution:} Rescue and retreat with his remains (Boons +1, but a ghost follows)

\textbf{Renunciation:} Leave without honoring him (Diamond: debt to the dead)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Comrade's Shield (once: +1 defense)}
\burden{Ghost follower if you retreat; Ledger - Helped: Thane if you stay}
\textbf{Shard Kit:} Fallen comrade's shield grants +1d to loyalty tests later; the shield bears Utaran military honors
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored his memory; his service record connects to the Siege of Ecktoria
\textbf{Name Ledger Update:} Harmed: Thane Ironheart • Helped: [ ] • Unresolved: [ ]

\entrytext{The sand shifts beneath your feet, whispering the names of the fallen. Banners above flutter like nooses, and the air tastes of old iron and regret. Among the shifting dunes, fragments of Utaran military equipment lie buried, silent witnesses to past battles.}

\exittext{The sand settles into a proper grave as Thane's name is spoken. His shield gleams once before going dark, and the battlefield falls silent. The distant echo of Utaran military bugles fades with honor.}
\end{floorbox}

% --- Floor 4 ---
\begin{floorbox}{4 --- The Rejected Love}
\memoryline{He spurned mortal affection to pursue immortality.}
\tonetags{contrition • night • garden • heartbreak}
\startposition{Dominant}
\floorclock{[4] --- Flowers close; heart slows; silver lilies wilt.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} Love was a distraction, he told himself---a weakness that would prevent him from achieving true greatness. The mortal woman who loved him was beautiful, kind, and patient, but he could not see past his own ambitions. When she offered her heart, he chose his research instead, never realizing that in rejecting her, he had already lost the very thing he sought to preserve. As the Utaran Empire crumbled around him, his pursuit of power over personal connection became emblematic of the empire's own failures.

\textbf{From the Harmed Party's Perspective:} She had given everything---her heart, her future, her dreams of a simple life together. In return, she received nothing but cold words and a slammed door. Justice would have been recognition---acknowledgment that love deserved more than scraps from his ambition's table. Many Utaran citizens faced similar heartbreak as their empire prioritized state needs over personal happiness.

\textbf{Penance Sought:} To acknowledge the love he rejected and the pain he caused. The lich recognizes that his personal rejection of love mirrored the empire's rejection of its people's needs.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Night garden of silver lilies; a heart in a glass cloche beats slowly; moonlight filters through twisted branches. The garden's design reflects Utaran aesthetic principles---symmetrical layouts with laurel-circle motifs and river-stone pathways.

\textbf{Anchors:} Glass cloche with beating heart, silver lilies arranged in Utaran noble patterns, love letters bearing Utaran postal markings, memory-tree with carved Utaran civic symbols

\textbf{Senses:} The scent of night-blooming flowers; the sound of a slowing heartbeat; moonlight that seems to watch; the distant echo of the Noon Bell that once marked time in Utaran gardens.

\textbf{Adversaries/Agents:} Wilted lilies that attack, memory echoes of romantic moments, guardian of the heart, the ghost of a Utaran garden keeper demanding proper respect for nature.
\end{gmnote}

\handles{Sway (sincere apology, DV 2) • Theft (return a token, DV 3) • Arcana (unbind cloche, DV 3) • Performance (dance the pledge back, DV 2)}

\textbf{Position Drift:} Returning the token raises Position next floor; breaking and burying drops it.

\textbf{When the Floor Clock Fills:} Flowers close; heart slows; silver lilies wilt and turn to thorns. A wilted lily reveals a Utaran edict restricting personal gatherings during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Flowers whisper your worst doubt. Among the whispers, echoes of Utaran propaganda about prioritizing state over self.}
    \sbitem{2}{The cloche draws breath from the room. The breath carries the scent of Utaran gardens that were destroyed during the Siege of Ecktoria.}
    \sbitem{3}{A hidden cache of love letters reveals the depth of feeling rejected. The letters reference social gatherings that were cancelled as Utaran resources were diverted to the war effort.}
    \sbitem{4+}{The guardian of the heart awakens, demanding satisfaction. The guardian wears the robes of a Utaran courtier who died when the empire's social structures collapsed.}
\end{description}

\deckdraw{Draw when the heart's beat slows or when the guardian awakens---omens of rejected love, debts of emotional harm, echoes of Utaran social collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{LYRA NIGHTINGALE}
\nameimportance{Restoring her full name acknowledges she was more than the wound she became. Her family name connects to Utaran nobility who lost their estates when the empire fell.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Palimpsest: Her name visible only when tears fall on the cloche, revealed beneath erased Utaran social invitations
    \item Emotional Key: Appears when a PC chooses compassion over ambition, echoing the Utaran values of personal honor and social grace
    \item Syllable Scavenger: Fragments of her name found in love letters bearing Utaran postal cancellations from the years before the Siege of Ecktoria
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Silver lilies arranged in her initials, with the arrangement matching Utaran noble family crests
    \item[Actionable Lead:] Theft reveals a hidden locket with her portrait, the locket bearing Utaran noble markings and a date from before the empire's collapse
    \item[Aha! Proof:] Sway work uncovers her true feelings in her own words, including references to social events cancelled due to Utaran wartime measures
\end{description}

\textbf{Extraction Test:} Sway or Theft to recover her token (on fail, the heart's beat weakens but reveals her middle name and family connections to Utaran refugee networks).

\textbf{Verification \& Use:} Speak her name while returning her token to the cloche; mispronunciations cause the lilies to wither (harm 1) and add a clock segment. Speaking her full name correctly causes the garden to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Return the token (gain Favor; +1d once later)

\textbf{Restitution:} Break and bury (end the beat; add Gloom tag next floor)

\textbf{Renunciation:} Promise to carry (take a vow-Condition to clear Gloom)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Lover's Token (once: +1d to sincere social actions)}
\burden{Gloom tag if you break it; Ledger - Helped: Lyra if you return it}
\textbf{Shard Kit:} Silver lily grants +1d to recognize true love later; the lily's stem bears Utaran noble markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her family's refugee story connects to Utaran's fall
\textbf{Name Ledger Update:} Harmed: Lyra Nightingale • Helped: [ ] • Unresolved: [ ]

\entrytext{Silver lilies bloom in impossible moonlight, their petals shimmering like stars. In a glass cloche at the garden's heart, something beats with slow, sad rhythm. The garden's paths are laid with river-stone in the style of Utaran noble estates, now overgrown and forgotten.}

\exittext{The lilies' glow softens as Lyra's name is spoken. The heart's beat steadies, and for a moment, you feel the warmth of love that was once offered. The distant echo of the Noon Bell fades with satisfaction, and the garden remembers its former beauty.}
\end{floorbox}

% --- Floor 5 ---
\begin{floorbox}{5 --- The Betrayed Mentor}
\memoryline{He stole forbidden pages.}
\tonetags{contrition • study • theft • hypocrisy}
\startposition{Dominant}
\floorclock{[5] --- Footnotes recite accusations; cane taps faster; secret pages flare.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} Knowledge was power, and power was the only thing that mattered. His mentor had hoarded forbidden texts, keeping them from those who might use them for good. He told himself he was liberating wisdom, making it available to those who deserved it. But the truth was simpler: he coveted what his teacher possessed and was willing to betray trust to obtain it. During the Siege of Ecktoria, such betrayals of trust would become common as desperate officials sought any advantage.

\textbf{From the Harmed Party's Perspective:} The forbidden pages were not his to give---he had sworn oaths to protect them, to use them only under strict conditions. When his student stole them, he felt not just the loss of precious knowledge but the deeper wound of betrayal by one he had trusted implicitly. Justice would have been acknowledgment of the theft and restoration of what was taken. Many Utaran scholars faced similar betrayals as their institutions collapsed.

\textbf{Penance Sought:} To acknowledge his betrayal and make amends for stealing his mentor's trust. The lich recognizes that his academic dishonesty contributed to the erosion of scholarly integrity during Utaran's final years.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Spiral study; every step recites footnotes; the mentor's cane taps from above; forbidden texts line the walls. The study's architecture reflects Utaran scholarly traditions---triple-course river-stone walls and laurel-circle ceiling medallions.

\textbf{Anchors:} Mentor's cane bearing Utaran academic seals, forbidden pages with Utaran library stamps, study ledger with Utaran administrative numbering, accusation scrolls bearing Utaran legal formulas

\textbf{Senses:} The sound of whispered footnotes; the tap-tap of the cane growing closer; the musty scent of ancient parchment; the distant echo of the Noon Bell that once marked time in Utaran academies.

\textbf{Adversaries/Agents:} Marginalia-imps, secret page guardians, mentor's echo, the ghost of a Utaran librarian demanding proper handling of knowledge.
\end{gmnote}

\handles{Deception (stage a confession, DV 3) • Tinker (swap facsimile pages, DV 2) • Arcana (trace original ink, DV 3) • Skirmish (duel the cane, DV 2)}

\textbf{Position Drift:} Returning pages raises Position next floor; keeping knowledge drops it.

\textbf{When the Floor Clock Fills:} Footnotes recite accusations louder; cane taps faster; secret pages flare with dangerous energy. A secret page reveals Utaran academic regulations on the proper handling of forbidden knowledge.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{A marginalia-imp bites at evidence. The imp carries the scent of Utaran academic halls that were destroyed during the Siege of Ecktoria.}
    \sbitem{2}{Accusation scrolls multiply, creating confusion. The scrolls bear the official seals of Utaran academic institutions that no longer exist.}
    \sbitem{3}{A secret page flares, blinding. The page contains research that might have helped defend Ecktoria, had it been shared properly.}
    \sbitem{4+}{The mentor's echo arrives to confront the thief. The echo wears the robes of a Utaran scholar who died when the empire's libraries were destroyed.}
\end{description}

\deckdraw{Draw when the secret page flares or when the mentor's echo arrives---omens of academic betrayal, debts of stolen wisdom, echoes of Utaran intellectual collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{MASTER RHEM, KEEPER OF FORBIDDEN LORE}
\nameimportance{Acknowledging his full title recognizes the respect and trust that was betrayed. His position as Keeper connects to Utaran academic traditions, and his death during the empire's collapse represents the loss of institutional knowledge.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of book titles spell his name, with titles bearing Utaran library catalog numbers
    \item Palimpsest: His name visible only under the light of forbidden knowledge, revealed beneath erased Utaran academic credentials
    \item Ledger Lock: Must be written with a specific quill on a named page, the quill being one used in Utaran academic ceremonies
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Book spines arranged in his initials, with the arrangement matching Utaran academic honorifics
    \item[Actionable Lead:] Arcana reveals energy traces leading to his hidden study, traces that match patterns used in Utaran scholarly magic
    \item[Aha! Proof:] Tinker work uncovers his secret ledger with the full name, including his title as Keeper and references to Utaran academic protocols
\end{description}

\textbf{Extraction Test:} Arcana or Deception to trace his contributions (on fail, the forbidden pages become more dangerous but reveal his middle name and academic credentials).

\textbf{Verification \& Use:} Speak his full name while returning the pages to his study; mispronunciations cause the books to rebel (harm 1) and add a clock segment. Speaking his full title correctly causes the study to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Return pages (Position +1)

\textbf{Restitution:} Keep knowledge (gain Technique, but GM banks a Debt)

\textbf{Renunciation:} Expose hypocrisy (mentor-echo breaks; Consequences Deck draw now)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Scholar's Knowledge (once: +1d to forbidden lore)}
\burden{Debt if you keep knowledge; Ledger - Helped: Master Rhem if you return pages}
\textbf{Shard Kit:} Mentor's quill grants +1d to recognize hypocrisy later; the quill bears Utaran academic markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored his name; his death during Utaran's collapse represents institutional loss
\textbf{Name Ledger Update:} Harmed: Master Rhem • Helped: [ ] • Unresolved: [ ]

\entrytext{The spiral study whispers with the voices of a thousand footnotes, each one an accusation. Above, the tap-tap of a cane grows steadily closer, echoing the rhythm of Utaran academic ceremonies now lost to time. The walls are lined with triple-course river-stone, bearing the scars of past conflicts.}

\exittext{The footnotes fall silent as Master Rhem's name is spoken. The forbidden pages settle into their proper places, and the study finds peace. The distant echo of the Noon Bell fades with satisfaction, and the study remembers its former purpose as a place of learning.}
\end{floorbox}

% --- Floor 6 ---
\begin{floorbox}{6 --- The Children of the Plague-House}
\memoryline{He quarantined too harshly, dooming survivors.}
\tonetags{contrition • infirmary • plague • mercy}
\startposition{Dominant}
\floorclock{[6] --- Lullaby grows discordant; vents cough worse fumes; time loss imminent.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} Containment was paramount---any breach could mean the spread of pestilence beyond the house's walls. He had seen what unchecked disease could do, how it could ravage entire communities. The harsh quarantine was necessary, he told himself, even if it meant some would perish who might otherwise have lived. But in his zeal for protection, he had forgotten that mercy was also a form of wisdom. During the Siege of Ecktoria, such harsh measures would become common as Utaran authorities struggled with plague outbreaks.

\textbf{From the Harmed Party's Perspective:} They were not the disease---they were children who had lost their families to it, who had found some measure of safety within the house's walls. When the quarantine was sealed too tightly, when no one came to tend their needs, they faced a slower death from neglect. Justice would have been care---someone to feed them, comfort them, give them a chance. Many Utaran children faced similar neglect as their empire's social services collapsed.

\textbf{Penance Sought:} To acknowledge the mercy he failed to show and the lives lost to his harsh judgment. The lich recognizes that his medical failures mirrored the empire's abandonment of its most vulnerable citizens.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Boarded infirmary, lullaby through vents, medical supplies scattered and decaying. The infirmary's design reflects Utaran medical standards---river-stone walls with laurel-circle ventilation grates and triple-course construction.

\textbf{Anchors:} Quarantine ledger with Utaran administrative markings, medical supplies bearing Utaran health ministry stamps, children's drawings showing Utaran civic symbols, lullaby box with Utaran musical notation

\textbf{Senses:} The scent of old medicine and decay; the sound of a discordant lullaby; the feel of cold, sterile air; the distant echo of the Noon Bell that once marked time in Utaran hospitals.

\textbf{Adversaries/Agents:} Coughing echoes, plague spirits, time-loss phantoms, the ghost of a Utaran medical administrator demanding proper care protocols.
\end{gmnote}

\handles{Medicine (triage echoes, DV 2) • Insight (weigh utilitarian cost, DV 3) • Rites (quiet passing, DV 2) • Move (thread contagion wards, DV 3)}

\textbf{Position Drift:} Showing mercy raises Position next floor; iron law drops it.

\textbf{When the Floor Clock Fills:} Lullaby grows discordant; vents cough worse fumes; time loss becomes imminent. A medical supply crate reveals Utaran health ministry regulations on proper quarantine procedures.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Cough spreads to a PC (Fatigue). The cough carries the distinctive sound of Utaran plague victims from the Siege of Ecktoria.}
    \sbitem{2}{Medical supplies prove inadequate. Among the supplies are Utaran emergency rations that were distributed during wartime health crises.}
    \sbitem{3}{A screaming vent opens---time loss. Through the vent, echoes of Utaran medical personnel who died during plague outbreaks can be heard.}
    \sbitem{4+}{The spirits of the children demand justice. The spirits wear the clothes of Utaran children who died when the empire's social services collapsed.}
\end{description}

\deckdraw{Draw when the cough spreads or when the children's spirits rise---omens of medical failure, debts of unfulfilled care, echoes of Utaran social collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE CHILDREN OF THE WHITE ROSE INFIRMARY}
\nameimportance{They were more than victims---they were individuals who deserved better. The infirmary's name connects to Utaran medical traditions, and their deaths represent the empire's failure to protect its young.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Syllable Scavenger: Fragments of their collective name found in children's drawings showing Utaran civic symbols
    \item Palimpsest: Their name visible only when tears fall on the lullaby box, revealed beneath erased Utaran medical records
    \item Emotional Key: Appears when a PC chooses care over containment, echoing the Utaran values of protecting the vulnerable
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Children's drawings spell out "White Rose" in the style of Utaran noble family crests
    \item[Actionable Lead:] Medicine work reveals their individual names in medical records bearing Utaran health ministry seals
    \item[Aha! Proof:] Insight uncovers their true story in hidden journals, including references to Utaran social services that were overwhelmed during wartime
\end{description}

\textbf{Extraction Test:} Medicine or Insight to uncover their full story (on fail, the lullaby grows more discordant but reveals more names and connections to Utaran refugee networks).

\textbf{Verification \& Use:} Speak their collective name while tending to their echoes; mispronunciations cause the plague to worsen (harm 1) and add a clock segment. Speaking their full name correctly causes the infirmary to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Mercy (gain Blessing of Rest)

\textbf{Iron Law:} Desperate next floor; but gain Efficiency boon later

\textbf{Risk It All:} Save two echoes; take Debt
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Blessing of Rest (once: clear Fatigue)}
\burden{Debt if you risk it all; Ledger - Helped: Children if you show mercy}
\textbf{Shard Kit:} Lullaby box grants +1d to recognize suffering later; the box bears Utaran medical markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their memory; their deaths during Utaran's collapse represent social failure
\textbf{Name Ledger Update:} Harmed: Children of the White Rose Infirmary • Helped: [ ] • Unresolved: [ ]

\entrytext{The infirmary reeks of old medicine and something worse. Through the vents comes a lullaby, discordant and wrong, while shadows move behind boarded windows. The walls are constructed of triple-course river-stone, bearing the official seals of Utaran health ministry that no longer function.}

\exittext{The lullaby softens into a proper tune as the children's names are spoken. Their echoes fade with smiles, and the infirmary finds peace. The distant echo of the Noon Bell fades with satisfaction, and the infirmary remembers its former purpose as a place of healing.}
\end{floorbox}

% --- Floor 7 ---
\begin{floorbox}{7 --- The Crown Informant}
\memoryline{He traded a friend to the authorities.}
\tonetags{contrition • tea-house • masks • betrayal}
\startposition{Dominant}
\floorclock{[4] --- Masks tilt accusingly; whispers grow louder; sealed writ lands.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The choice seemed clear at the time---protect himself and his research by eliminating a threat. His friend had stumbled upon dangerous knowledge, knowledge that could destroy everything he had built. He told himself it was necessary, that the greater good required sacrifice. But in the end, it was fear that drove him to betray the trust of one who had stood by him through everything. During the Siege of Ecktoria, such betrayals would become common as Utaran authorities turned citizen against citizen.

\textbf{From the Harmed Party's Perspective:} They had been friends for years, sharing secrets, dreams, and dangers. When the authorities came, they never expected their friend to be the one to open the door. Justice would have been loyalty---standing together against the storm, finding another way, anything but betrayal. Many Utaran citizens faced similar betrayals as their empire's security apparatus turned inward.

\textbf{Penance Sought:} To acknowledge the betrayal and the trust that was broken. The lich recognizes that his personal betrayal mirrored the empire's betrayal of its citizens' trust.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Tea-house murmurs; lacquered masks tilt toward you; shadows dance between tables. The tea-house's design reflects Utaran social traditions---laurel-circle decorative motifs and river-stone construction with triple-course walls.

\textbf{Anchors:} Lacquered masks bearing Utaran noble markings, friendship tokens with Utaran civic symbols, sealed writ bearing Utaran legal seals, tea service with Utaran artisan markings

\textbf{Senses:} The scent of jasmine tea and betrayal; the sound of whispered accusations; the feel of watching eyes; the distant echo of the Noon Bell that once marked time in Utaran social gatherings.

\textbf{Adversaries/Agents:} Mask-enforcers, friend's echo, informant spirits, the ghost of a Utaran legal administrator demanding proper procedure.
\end{gmnote}

\handles{Deception (cover story, DV 3) • Sway (reframe as lesser betrayal, DV 2) • Streetwise (flip the informant chain, DV 3) • Skirmish (mask-enforcers, DV 2)}

\textbf{Position Drift:} Confessing raises Position next floor; doubling down drops it.

\textbf{When the Floor Clock Fills:} Masks tilt accusingly; whispers grow louder; sealed writ lands with finality. The writ bears Utaran legal formulas for denunciation procedures used during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Friend-echo recognizes you. The echo wears the clothes of a Utaran citizen who was denounced during the Siege of Ecktoria.}
    \sbitem{2}{Tea-house patrons turn hostile. Among the patrons are echoes of Utaran officials who used such establishments for surveillance.}
    \sbitem{3}{A sealed writ lands---law binds choices. The writ bears the official seal of a Utaran security ministry that operated during wartime.}
    \sbitem{4+}{The friend's spirit arrives to confront the betrayer. The spirit wears the uniform of a Utaran citizen who died in a security prison.}
\end{description}

\deckdraw{Draw when the friend's echo recognizes you or when the sealed writ lands---omens of betrayed trust, debts of broken friendship, echoes of Utaran security collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{KIRA THORNWALKER, FRIEND BETRAYED}
\nameimportance{Restoring her full name acknowledges she was more than the wound she became. Her family name connects to Utaran nobility who lost everything when the empire fell.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of tea blends spell her name, with blends bearing Utaran trade guild markings
    \item Palimpsest: Her name visible only when tea is spilled on friendship tokens, revealed beneath erased Utaran social invitations
    \item Emotional Key: Appears when a PC chooses honesty over self-preservation, echoing the Utaran values of personal honor
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Tea blends arranged in her initials, with the arrangement matching Utaran noble family crests
    \item[Actionable Lead:] Sway reveals her true feelings in hidden letters, including references to Utaran social gatherings that were cancelled during wartime
    \item[Aha! Proof:] Streetwise work uncovers her full story in informant records bearing Utaran security ministry seals
\end{description}

\textbf{Extraction Test:} Sway or Streetwise to uncover her full story (on fail, the tea-house grows more hostile but reveals her middle name and family connections to Utaran refugee networks).

\textbf{Verification \& Use:} Speak her name while returning her friendship token; mispronunciations cause the masks to turn hostile (harm 1) and add a clock segment. Speaking her full name correctly causes the tea-house to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Confess (gain Forgiveness token; spend later to ease a social DV)

\textbf{Restitution:} Double Down (gain leverage; group marks Stain)

\textbf{Renunciation:} Cut the Chain (burn the writ; Diamond omen tracks you)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Forgiveness Token (once: ease social DV)}
\burden{Stain if you double down; Ledger - Helped: Kira if you confess}
\textbf{Shard Kit:} Friendship token grants +1d to recognize betrayal later; the token bears Utaran noble markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her family's refugee story connects to Utaran's fall
\textbf{Name Ledger Update:} Harmed: Kira Thornwalker • Helped: [ ] • Unresolved: [ ]

\entrytext{The tea-house murmurs with a thousand whispered conversations, but all eyes seem to find you. Lacquered masks tilt in your direction, and the scent of jasmine grows sharp with accusation. The establishment's decor features laurel-circle motifs and river-stone construction, reminiscent of Utaran social venues that were shuttered during the empire's final years.}

\exittext{The masks settle as Kira's name is spoken. Her friendship token warms briefly before going cold, and the tea-house finds peace. The distant echo of the Noon Bell fades with satisfaction, and the establishment remembers its former purpose as a place of honest fellowship.}
\end{floorbox}

% --- Floor 8 ---
\begin{floorbox}{8 --- The Oath of the Doorwarden}
\memoryline{He failed a sacred watch.}
\tonetags{contrition • temple • door • vigil}
\startposition{Dominant}
\floorclock{[5] --- Hinge screams; pilgrim-echo pushes through; breach-creature stirs.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The sacred trust placed in him was immense---a door that kept the darkness at bay, that protected the innocent who sought sanctuary. But he had grown weary, complacent in his duty. When the moment of greatest need came, he faltered, his attention wandering just long enough for the breach to occur. The consequences were catastrophic, and he knew that no amount of penance could truly make amends. During the Siege of Ecktoria, such failures of duty would prove devastating as defenders faltered under pressure.

\textbf{From the Harmed Party's Perspective:} They had come seeking sanctuary, trusting in the ancient oaths that protected this place. When the door failed, when the guardian fell asleep at his post, they were left to face horrors beyond imagination. Justice would have been vigilance---unwavering attention to duty, no matter how long or boring the watch. Many Utaran citizens faced similar failures as their empire's defenders grew weary during prolonged conflicts.

\textbf{Penance Sought:} To acknowledge his failure and stand a proper watch. The lich recognizes that his personal failure of duty mirrored the empire's failure to protect its people.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Temple gate, seven locks, one missing tooth; ancient wards flicker and fade. The temple's architecture reflects Utaran religious traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Temple gate bearing Utaran religious seals, seven locks with Utaran administrative numbering, missing tooth that once bore Utaran military markings, guardian's helm with Utaran honorifics, sacred oil in Utaran ceremonial vessels

\textbf{Senses:} The scent of old incense and neglect; the sound of a screaming hinge; the feel of darkness pressing at the edges; the distant echo of the Noon Bell that once marked time in Utaran temples.

\textbf{Adversaries/Agents:} Breach-creatures, pilgrim-echoes, failing wards, the ghost of a Utaran religious administrator demanding proper vigilance.
\end{gmnote}

\handles{Tinker (forge tooth, DV 3) • Rites (reconsecrate, DV 2) • Endure (hold the door, DV 3) • Shoot (pin the breach-creature, DV 2)}

\textbf{Position Drift:} Standing the watch raises Position next floor; opening to all drops it.

\textbf{When the Floor Clock Fills:} The hinge screams; pilgrim-echo pushes through; breach-creature stirs from its slumber. A pilgrim's pack reveals Utaran religious texts that were destroyed during wartime conflicts.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{The hinge screams with increasing intensity. The sound echoes the screams of Utaran civilians who died when their sanctuaries were breached during the Siege of Ecktoria.}
    \sbitem{2}{Sacred oil runs low. Among the remaining oil are reserves that were set aside during Utaran religious festivals that were cancelled during wartime.}
    \sbitem{3}{A pilgrim-echo tries to push through. The echo wears the clothes of a Utaran pilgrim who died when religious travel was restricted during wartime emergencies.}
    \sbitem{4+}{The breach-creature fully awakens. The creature bears the markings of Utaran military experiments that went wrong during the empire's final campaigns.}
\end{description}

\deckdraw{Draw when the breach-creature stirs or when pilgrim-echoes push through---omens of failed duty, debts of broken oaths, echoes of Utaran religious collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE OATH CALLED THRESHOLD}
\nameimportance{The oath itself was the true guardian, and by failing it, he failed everything it protected. The oath's name connects to Utaran religious traditions, and its breaking represents the empire's abandonment of its spiritual foundations.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of the seven locks spell the oath, with locks bearing Utaran administrative numbering systems
    \item Palimpsest: The oath visible only when sacred oil is applied to the missing tooth, revealed beneath erased Utaran religious texts
    \item Emotional Key: Appears when a PC chooses duty over convenience, echoing the Utaran values of honor and service
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] The seven locks arranged in the shape of an oath, with the arrangement matching Utaran religious symbols
    \item[Actionable Lead:] Rites reveal the missing words of the sacred vow, words that reference Utaran civic duties and religious obligations
    \item[Aha! Proof:] Tinker work uncovers the full text in ancient records, including references to Utaran religious ceremonies that were suspended during wartime
\end{description}

\textbf{Extraction Test:} Rites or Tinker to restore the full oath (on fail, the breach grows wider but reveals more of the text and connections to Utaran religious collapse).

\textbf{Verification \& Use:} Speak the full oath while placing the forged tooth; mispronunciations cause the door to weaken (harm 1) and add a clock segment. Speaking the full oath correctly causes the temple to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Stand the Watch (Position +1 next two floors)

\textbf{Restitution:} Open to All (gain Popular Favor; Club draw now)

\textbf{Renunciation:} Bar Forever (close one shortcut; reduce later DV by 1 once)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Guardian's Vigil (once: +1d to guard duties)}
\burden{Popular Favor if you open the door; Ledger - Helped: Pilgrims if you stand watch}
\textbf{Shard Kit:} Guardian's helm grants +1d to recognize duty later; the helm bears Utaran religious markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored the oath; its breaking represents Utaran's religious collapse
\textbf{Name Ledger Update:} Harmed: Pilgrims seeking sanctuary • Helped: [ ] • Unresolved: [ ]

\entrytext{The temple gate stands before you, its seven locks gleaming with ancient power. But one tooth is missing from its great mouth, and through the gap, you can feel something dark stirring. The gate's construction features triple-course river-stone and laurel-circle motifs, reminiscent of Utaran religious architecture that was destroyed during wartime conflicts.}

\exittext{The gate's locks click into place as the oath is spoken. The missing tooth settles into position, and the temple finds peace. The distant echo of the Noon Bell fades with satisfaction, and the temple remembers its former purpose as a place of sanctuary.}
\end{floorbox}

% --- Floor 9 ---
\begin{floorbox}{9 --- The Scholar's Duel}
\memoryline{He discredited a rival through lies.}
\tonetags{contrition • lecture-hall • chalk • falsehood}
\startposition{Dominant}
\floorclock{[4] --- Bell tolls on falsehoods; note-mites chew proofs; reputation crumbles.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} Academic reputation was everything---without it, his research would be dismissed, his theories ignored. When his rival presented compelling evidence that threatened to overshadow his own work, he knew he had to act. The lies came easily, spun from half-truths and careful omissions. He told himself it was a necessary evil, that the greater good of his research justified the deception. But in the end, he had only proven that he was unworthy of the very reputation he sought to protect. During the Siege of Ecktoria, such academic dishonesty would become common as desperate scholars sought any advantage.

\textbf{From the Harmed Party's Perspective:} Years of research, countless hours of careful study, all destroyed by a rival's malicious lies. The academic community turned their back, their funding was cut, and their career was left in ruins. Justice would have been truth---the chance to present their evidence fairly, to let the merit of their work speak for itself. Many Utaran scholars faced similar betrayals as their empire's intellectual community collapsed.

\textbf{Penance Sought:} To acknowledge his lies and restore his rival's reputation. The lich recognizes that his academic dishonesty contributed to the erosion of scholarly integrity during Utaran's final years.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Lecture-hall, chalk specters, a bell that tolls on falsehoods; academic papers scattered and torn. The lecture-hall's design reflects Utaran academic traditions---triple-course river-stone walls with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Chalkboard bearing Utaran academic formulas, bell of truth with Utaran educational seals, rival's research with Utaran library stamps, discrediting documents bearing Utaran administrative markings

\textbf{Senses:} The scent of chalk dust and shame; the sound of the truth bell tolling; the feel of academic judgment; the distant echo of the Noon Bell that once marked time in Utaran academies.

\textbf{Adversaries/Agents:} Chalk specters, note-mites, academic censors, the ghost of a Utaran academic administrator demanding proper scholarly conduct.
\end{gmnote}

\handles{Evidence (reconstruct data, DV 3) • Sway (track to apology, DV 2) • Deception (plant a correction, DV 3) • Skirmish (ink-spirits, DV 2)}

\textbf{Position Drift:} Publishing retraction raises Position next floor; winning again drops it.

\textbf{When the Floor Clock Fills:} The bell tolls wrong---everyone loses a turn; note-mites devour more proofs. A torn paper reveals Utaran academic regulations on proper scholarly conduct during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Note-mites chew proofs. Among the destroyed papers are Utaran academic research that might have helped during the Siege of Ecktoria.}
    \sbitem{2}{Chalk specters rewrite history. The rewritten history omits the contributions of Utaran scholars who died during the empire's collapse.}
    \sbitem{3}{The bell tolls wrong---everyone loses a turn. The discordant toll echoes the chaos that overtook Utaran academic institutions during wartime.}
    \sbitem{4+}{The rival's spirit arrives to demand justice. The spirit wears the robes of a Utaran scholar who died when the empire's libraries were destroyed.}
\end{description}

\deckdraw{Draw when the bell tolls wrong or when the rival's spirit arrives---omens of academic dishonesty, debts of ruined careers, echoes of Utaran intellectual collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{DR. ELYSIA GREENWELL, SCHOLAR BETRAYED}
\nameimportance{Restoring her full name acknowledges she was more than the wound she became. Her academic title connects to Utaran scholarly traditions, and her research represented knowledge lost when the empire fell.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of academic titles spell her name, with titles bearing Utaran university credentials
    \item Palimpsest: Her name visible only under the light of the truth bell, revealed beneath erased Utaran academic publications
    \item Emotional Key: Appears when a PC chooses honesty over ambition, echoing the Utaran values of scholarly integrity
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Academic papers arranged in her initials, with the arrangement matching Utaran academic honorifics
    \item[Actionable Lead:] Evidence reveals her true contributions in hidden files, including research on defensive magic that might have helped during the Siege of Ecktoria
    \item[Aha! Proof:] Sway work uncovers her vindication in peer reviews bearing Utaran academic seals
\end{description}

\textbf{Extraction Test:} Evidence or Sway to restore her reputation (on fail, more proofs are destroyed but reveal her middle name and academic credentials).

\textbf{Verification \& Use:} Speak her name while posting a retraction; mispronunciations cause the chalk specters to turn hostile (harm 1) and add a clock segment. Speaking her full name correctly causes the lecture-hall to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Publish Retraction (Boons +1)

\textbf{Restitution:} Win Again (gain Repute, but a critic haunt follows)

\textbf{Renunciation:} Bind the Bell (stabilize truth; Diamond licenses your re-entry)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Academic Integrity (once: +1d to scholarly honesty)}
\burden{Critic haunt if you win again; Ledger - Helped: Dr. Greenwell if you retract}
\textbf{Shard Kit:} Truth bell fragment grants +1d to detect lies later; the fragment bears Utaran academic markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her research represents knowledge lost during Utaran's collapse
\textbf{Name Ledger Update:} Harmed: Dr. Elysia Greenwell • Helped: [ ] • Unresolved: [ ]

\entrytext{The lecture-hall echoes with the whispers of a thousand debates, but all eyes are on the bell that tolls with increasing urgency. Chalk specters hover over torn papers, and the air hums with the weight of academic judgment. The hall's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran university buildings that were destroyed during wartime conflicts.}

\exittext{The bell's toll steadies as Dr. Greenwell's name is spoken. Her research papers flutter into proper order, and the hall finds peace. The distant echo of the Noon Bell fades with satisfaction, and the lecture-hall remembers its former purpose as a place of learning.}
\end{floorbox}

% --- Floor 10 ---
\begin{floorbox}{10 --- The Siege of Hearth-Row}
\memoryline{He sold a neighborhood to fund research.}
\tonetags{contrition • siege • hearth • betrayal}
\startposition{Dominant}
\floorclock{[6] --- Roof collapse; hostage ledger appears; house-gods glare.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The research required resources---rare components, specialized equipment, assistants who understood the work. When his funds ran low, he saw an opportunity in the struggling neighborhood of Hearth-Row. The residents were desperate, willing to accept any offer that promised stability. He told himself he was helping them, giving them a better future, but the truth was that he was using their desperation to fund his own ambitions. During the Siege of Ecktoria, such exploitation of the desperate would become common as officials sought resources by any means.

\textbf{From the Harmed Party's Perspective:} They had lived in Hearth-Row for generations, building lives, families, and memories within its narrow lanes. When the researcher came with his offers, they thought they were securing a better future. Instead, they found themselves displaced, their homes sold out from under them to fund experiments they never agreed to. Justice would have been choice---the right to decide their own fate, not have it sold behind their backs. Many Utaran citizens faced similar displacements as their empire requisitioned property for wartime needs.

\textbf{Penance Sought:} To acknowledge his betrayal and the homes he destroyed. The lich recognizes that his exploitation mirrored the empire's exploitation of its citizens during desperate times.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Narrow lanes; house-gods glare from soot-stained walls; eviction notices flutter in the wind. The neighborhood's architecture reflects Utaran urban planning---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} House-gods bearing Utaran religious markings, eviction notices with Utaran legal seals, hostage ledger with Utaran administrative numbering, neighborhood map showing Utaran property boundaries

\textbf{Senses:} The scent of burning wood and displacement; the sound of weeping from empty homes; the feel of betrayed trust; the distant echo of the Noon Bell that once marked time in Utaran neighborhoods.

\textbf{Adversaries/Agents:} House-god spirits, eviction enforcers, displaced residents, the ghost of a Utaran property administrator demanding proper compensation.
\end{gmnote}

\handles{Command (evacuate echoes, DV 3) • Trade (find funds, DV 2) • Skirmish (block enforcers, DV 3) • Rites (invoke hearth-right, DV 2)}

\textbf{Position Drift:} Buying back time raises Position next floor; naming names drops it.

\textbf{When the Floor Clock Fills:} Roof collapse; hostage ledger appears; house-gods glare with increasing fury. The ledger reveals Utaran property requisition records from the Siege of Ecktoria.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Eviction notices multiply. Among the notices are forms bearing Utaran legal seals for wartime property seizures.}
    \sbitem{2}{Roof collapse blocks a path. The collapsed roof reveals Utaran building permits that were never properly processed during wartime emergencies.}
    \sbitem{3}{A hostage ledger appears. The ledger contains names of Utaran citizens who were displaced during the Siege of Ecktoria to make room for military installations.}
    \sbitem{4+}{The house-gods rise to defend their domain. The house-gods bear the markings of Utaran religious traditions that were suppressed during wartime.}
\end{description}

\deckdraw{Draw when the roof collapses or when the house-gods rise---omens of community betrayal, debts of broken homes, echoes of Utaran urban collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{HEARTH-ROW, THE BETRAYED NEIGHBORHOOD}
\nameimportance{The neighborhood itself was the victim, not just its individual residents. Hearth-Row's name connects to Utaran urban traditions, and its destruction represents the empire's betrayal of its citizens' homes.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of street names spell "Hearth-Row", with names bearing Utaran municipal numbering systems
    \item Palimpsest: The name visible only when soot is cleared from house-god walls, revealed beneath erased Utaran property records
    \item Emotional Key: Appears when a PC chooses community over individual gain, echoing the Utaran values of civic responsibility
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Street signs arranged to spell "Hearth-Row", with the arrangement matching Utaran municipal signage
    \item[Actionable Lead:] Trade reveals the full extent of the betrayal in financial records bearing Utaran treasury seals
    \item[Aha! Proof:] Command work uncovers the neighborhood's true history from survivor accounts, including references to Utaran wartime housing policies
\end{description}

\textbf{Extraction Test:} Trade or Command to uncover the full story (on fail, more homes are lost but reveal more details and connections to Utaran refugee networks).

\textbf{Verification \& Use:} Speak the neighborhood's name while invoking hearth-right; mispronunciations cause the house-gods to turn hostile (harm 1) and add a clock segment. Speaking the full name correctly causes the neighborhood to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Buy Back Time (lose coin/asset; +2 segments cleared)

\textbf{Restitution:} Name Names (Position -1; gain Coin)

\textbf{Renunciation:} Break the Ledger (boon: Hearth's Favor; draw Spade omen)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Hearth's Favor (once: +1d to community actions)}
\burden{Position -1 if you name names; Ledger - Helped: Hearth-Row if you buy back time}
\textbf{Shard Kit:} House-god fragment grants +1d to recognize community needs later; the fragment bears Utaran religious markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their name; the neighborhood's destruction represents Utaran's urban collapse
\textbf{Name Ledger Update:} Harmed: Hearth-Row residents • Helped: [ ] • Unresolved: [ ]

\entrytext{The narrow lanes of Hearth-Row stretch before you, lined with houses whose windows glow with the light of displaced souls. House-gods glare from soot-stained walls, and eviction notices flutter like accusing fingers. The neighborhood's construction features triple-course river-stone and laurel-circle motifs, reminiscent of Utaran urban districts that were requisitioned during wartime emergencies.}

\exittext{The house-gods' glares soften as Hearth-Row's name is spoken. The eviction notices burn away, and the neighborhood finds peace. The distant echo of the Noon Bell fades with satisfaction, and the district remembers its former purpose as a place of community.}
\end{floorbox}

% --- Floor 11 ---
\begin{floorbox}{11 --- The Pact of the Dry Well}
\memoryline{He broke a village's water-pact.}
\tonetags{contrition • drought • well • thirst}
\startposition{Dominant}
\floorclock{[5] --- Cracks spread underfoot; thirst-curse licks lips; spirit sighs below.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The village's water-pact was simple---regular offerings in exchange for a steady flow from the ancient well. But his research required rare components that grew only in the deepest caverns, beneath the village's sacred spring. He told himself the disruption would be temporary, that he could restore the flow once his work was complete. But the well's guardian was ancient and unforgiving, and the pact, once broken, could not be easily mended. During the Siege of Ecktoria, such breaking of sacred agreements would become common as desperate officials sought resources by any means.

\textbf{From the Harmed Party's Perspective:} The well had provided for generations, its waters sustaining crops, livestock, and families through countless seasons. When it suddenly ran dry, when the promised flow became a trickle and then nothing, the village faced starvation and despair. Justice would have been respect---honoring the ancient agreement, finding another way rather than breaking a sacred trust. Many Utaran villages faced similar betrayals as their empire's officials requisitioned resources for wartime needs.

\textbf{Penance Sought:} To acknowledge his broken promise and restore the village's water-source. The lich recognizes that his betrayal of sacred trust mirrored the empire's betrayal of its citizens' fundamental needs.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Salt rings mark where water once flowed; hung jars remain empty; something sighs from the dry well. The area's design reflects Utaran rural traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Dry well bearing Utaran municipal markings, salt rings arranged in Utaran sacred patterns, offering jars with Utaran religious symbols, water-pact scroll with Utaran legal seals, spirit's shrine with Utaran religious markings

\textbf{Senses:} The taste of dust and salt; the sound of distant sighing; the feel of creeping thirst; the distant echo of the Noon Bell that once marked time in Utaran rural communities.

\textbf{Adversaries/Agents:} Thirst-curse spirits, cracked earth elementals, village elders' echoes, the ghost of a Utaran water administrator demanding proper respect for natural resources.
\end{gmnote}

\handles{Rites (appease spirit, DV 2) • Survival (re-route flow, DV 3) • Tinker (raise bucket engine, DV 3) • Sway (mediate terms, DV 2)}

\textbf{Position Drift:} Patching the pact raises Position next floor; taking the depths drops it.

\textbf{When the Floor Clock Fills:} Cracks spread underfoot; thirst-curse licks your lips; the spirit's sigh grows into a roar. A cracked stone reveals Utaran water management records from the Siege of Ecktoria.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Cracks spread underfoot, threatening stability. The cracks follow patterns used in Utaran engineering projects that were abandoned during wartime.}
    \sbitem{2}{A hidden spring is discovered but guarded. The spring bears the markings of Utaran water rights that were suspended during wartime emergencies.}
    \sbitem{3}{A thirst-curse licks your lips, draining moisture. The curse carries the essence of Utaran drought victims from the Siege of Ecktoria.}
    \sbitem{4+}{The well-spirit fully manifests, demanding restitution. The spirit bears the markings of Utaran nature deities that were suppressed during wartime.}
\end{description}

\deckdraw{Draw when the thirst-curse manifests or when the well-spirit appears---omens of broken promises, debts of environmental harm, echoes of Utaran rural collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{NIMUE, GUARDIAN OF THE SACRED WELL}
\nameimportance{The spirit had a name and identity beyond being merely a resource to be exploited. Her name connects to Utaran religious traditions, and her betrayal represents the empire's abandonment of its spiritual foundations.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Palimpsest: Her name visible only when water is poured on ancient shrine, revealed beneath erased Utaran religious texts
    \item Syllable Scavenger: Fragments of her name found in village folklore bearing Utaran cultural markings
    \item Emotional Key: Appears when a PC chooses restoration over exploitation, echoing the Utaran values of respect for nature
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Village folklore mentions "Nimue's tears", with the folklore bearing Utaran oral tradition markers
    \item[Actionable Lead:] Rites reveal her true name in ritual texts bearing Utaran religious seals
    \item[Aha! Proof:] Survival work uncovers her shrine with the full name, including references to Utaran water management practices that were abandoned during wartime
\end{description}

\textbf{Extraction Test:} Rites or Survival to appease her properly (on fail, the thirst worsens but reveals her title and connections to Utaran religious collapse).

\textbf{Verification \& Use:} Speak her full name while making proper offerings; mispronunciations cause the cracks to widen (harm 1) and add a clock segment. Speaking her full name correctly causes the area to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Patch the Pact (gain Token: Water-right)

\textbf{Restitution:} Take the Depths (immediate aid; mark Obligation-style Condition)

\textbf{Renunciation:} Walk Away (Desperate start next floor)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Water-Right Token (once: +1d to water-based Survival)}
\burden{Obligation-style Condition if you take the depths; Ledger - Helped: Village if you patch pact}
\textbf{Shard Kit:} Offering jar grants +1d to recognize sacred agreements later; the jar bears Utaran religious markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her betrayal represents Utaran's spiritual collapse
\textbf{Name Ledger Update:} Harmed: Village of the Sacred Well • Helped: [ ] • Unresolved: [ ]

\entrytext{Salt rings mark where water once flowed freely, and empty jars hang like accusations from wooden beams. From the depths of the dry well comes a sigh that speaks of ancient disappointment. The area's construction features triple-course river-stone and laurel-circle motifs, reminiscent of Utaran rural water management systems that were abandoned during wartime emergencies.}

\exittext{The salt rings glisten as Nimue's name is spoken. Water begins to seep from the well's mouth, and the air grows moist with promise. The distant echo of the Noon Bell fades with satisfaction, and the area remembers its former purpose as a place of sacred trust.}
\end{floorbox}

% --- Floor 12 ---
\begin{floorbox}{12 --- The Unburied}
\memoryline{He let enemies lie unburied to study them.}
\tonetags{contrition • bones • windmill • flies}
\startposition{Dominant}
\floorclock{[4] --- Bone-sail cleaves cover; skull remembers childhood; flies grow louder.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The dead held secrets---knowledge of anatomy, of the moment of death, of what lay beyond. He told himself his studies were for the greater good, that understanding death could lead to better preservation of life. But the truth was simpler: curiosity had overwhelmed respect, and he had left bodies to rot in the open, their bones picked clean by scavengers while he took notes and made observations. During the Siege of Ecktoria, such disrespect for the dead would become common as overwhelmed officials failed to maintain proper burial practices.

\textbf{From the Harmed Party's Perspective:} Death should bring peace, a final rest honored by proper burial. When their fallen were left to the elements, to be picked apart by carrion birds and stripped bare by wind and weather, it was a final indignity added to loss. Justice would have been dignity---respectful treatment of the dead, acknowledgment that even enemies deserved the peace of proper interment. Many Utaran soldiers faced similar indignities as their empire's burial services collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his disrespect and give the dead their proper rest. The lich recognizes that his failure to honor the dead mirrored the empire's failure to honor its fallen citizens.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Windmill of bones; chorused flies say your name; scattered skeletons tell stories of violence. The area's design reflects Utaran military traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Bone windmill bearing Utaran military markings, skeletal remains with Utaran identification tags, burial shrouds with Utaran religious symbols, death records with Utaran administrative numbering, flies' chorus echoing Utaran funeral dirges

\textbf{Senses:} The smell of decay and old death; the sound of bone rattling in the wind; the feel of being watched by the dead; the distant echo of the Noon Bell that once marked time in Utaran military cemeteries.

\textbf{Adversaries/Agents:} Bonewights, corpse-flies, angry spirits, memory-skulls, the ghost of a Utaran military chaplain demanding proper burial rites.
\end{gmnote}

\handles{Rites (burial, DV 2) • Arcana (sever tethers, DV 3) • Skirmish (put down bonewights, DV 2) • Endure (carry the weight, DV 3)}

\textbf{Position Drift:} Proper burial raises Position next floor; cataloguing drops it.

\textbf{When the Floor Clock Fills:} Bone-sail cleaves cover; skull remembers your childhood; flies grow into a maddening chorus. A bone fragment reveals Utaran military burial regulations that were suspended during the Siege of Ecktoria.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Bone-sail cleaves cover, revealing more remains. Among the remains are Utaran soldiers who died in the Siege of Ecktoria.}
    \sbitem{2}{A bone-sail cleaves through defenses. The sail bears the markings of Utaran military equipment that was requisitioned during wartime.}
    \sbitem{3}{A skull remembers your childhood, speaking forbidden truths. The skull wears the uniform of a Utaran civilian who died when the empire's social services collapsed.}
    \sbitem{4+}{The flies coalesce into a massive, angry swarm. The swarm carries the essence of Utaran plague victims from the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the bone-sail cuts through or when the flies swarm---omens of desecrated rest, debts of violated peace, echoes of Utaran military collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE UNNAMED DEAD OF THE BATTLE OF BLACKWATER FORD}
\nameimportance{Even enemies deserved recognition in death, not to be forgotten as mere specimens. Their collective name connects to Utaran military history, and their desecration represents the empire's abandonment of its fallen soldiers.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Syllable Scavenger: Fragments of their collective identity found on dog tags and personal effects bearing Utaran military serial numbers
    \item Palimpsest: Their names visible only when proper burial rites begin, revealed beneath erased Utaran military records
    \item Emotional Key: Appears when a PC chooses respect over knowledge, echoing the Utaran values of honoring the fallen
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Dog tags spell out unit designations, with the designations matching Utaran military unit numbering systems
    \item[Actionable Lead:] Rites reveal individual identities in death records bearing Utaran military seals
    \item[Aha! Proof:] Arcana work uncovers their full story in battlefield reports, including references to Utaran military units that were sacrificed during the Siege of Ecktoria
\end{description}

\textbf{Extraction Test:} Rites or Arcana to identify them properly (on fail, more remains are disturbed but reveal unit names and connections to Utaran military collapse).

\textbf{Verification \& Use:} Speak their collective name while performing burial rites; mispronunciations cause the bonewights to turn hostile (harm 1) and add a clock segment. Speaking their full name correctly causes the area to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Bury (Blessing vs Death magic)

\textbf{Restitution:} Catalogue (+1 effect vs undead later; gain Stain)

\textbf{Renunciation:} Confession (reduce next DV by 1 once)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Peaceful Rest Blessing (once: +1d vs death magic)}
\burden{Stain if you catalogue; Ledger - Helped: Dead if you bury them}
\textbf{Shard Kit:} Burial shroud grants +1d to recognize desecration later; the shroud bears Utaran religious markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their rest; their desecration represents Utaran's military collapse
\textbf{Name Ledger Update:} Harmed: Unnamed dead of Blackwater Ford • Helped: [ ] • Unresolved: [ ]

\entrytext{The windmill's sails are woven from bleached bones that click and clatter in the breeze. Flies buzz in a constant chorus, and scattered skeletons tell the story of a battle long past. The area's construction features triple-course river-stone and laurel-circle motifs, reminiscent of Utaran military installations that were abandoned during wartime conflicts.}

\exittext{The bone-windmill slows as the dead are named. The flies settle into respectful silence, and for a moment, the battlefield finds peace. The distant echo of the Noon Bell fades with satisfaction, and the area remembers its former purpose as a place of honor for the fallen.}
\end{floorbox}

% --- Floor 13 ---
\begin{floorbox}{13 --- The Rival's Heart}
\memoryline{He stole a rival's lover through glamour.}
\tonetags{contrition • ballroom • mirrors • glamour}
\startposition{Dominant}
\floorclock{[5] --- Floor becomes slick with spilled secrets; mirror eats dice; reflections turn hostile.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} Professional rivalry was one thing, but when his competitor found love and happiness, something primal stirred within him. He told himself he was simply demonstrating his superior magical prowess, that seducing away another's beloved was merely a test of skill. But the truth was uglier---he had been driven by jealousy, by the need to prove that he could take anything he desired, even another man's heart. During the Siege of Ecktoria, such personal betrayals would become common as social structures collapsed under wartime pressure.

\textbf{From the Harmed Party's Perspective:} Love was hard-won, built on trust, shared dreams, and countless small moments of connection. When it was stolen away by trickery and illusion, when the beloved was convinced to leave through magical manipulation, it felt like a theft of the soul. Justice would have been choice---the right to love freely, without enchantment's interference. Many Utaran citizens faced similar heartbreaks as their empire's social fabric unraveled during wartime.

\textbf{Penance Sought:} To acknowledge his manipulation and restore his rival's chance at happiness. The lich recognizes that his personal betrayal mirrored the empire's betrayal of its citizens' fundamental rights to choose their own paths.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Ballroom of mirrors; your reflections accuse; crystal chandeliers hang like frozen tears. The ballroom's design reflects Utaran aristocratic traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Mirrors bearing Utaran noble markings, glamour focus with Utaran magical symbols, love letters with Utaran postal cancellations, rival's portrait with Utaran artistic styles, heart pendant with Utaran noble crests

\textbf{Senses:} The scent of roses and regret; the sound of whispered accusations from reflections; the feel of being watched by multiple selves; the distant echo of the Noon Bell that once marked time in Utaran aristocratic gatherings.

\textbf{Adversaries/Agents:} Mirror-doppels, glamour spirits, jealous echoes, heart's guardian, the ghost of a Utaran social arbiter demanding proper conduct.
\end{gmnote}

\handles{Performance (dance the truth, DV 2) • Sway (public apology, DV 3) • Arcana (break glamour, DV 3) • Skirmish (mirror-doppel fights, DV 2)}

\textbf{Position Drift:} Undoing glamour raises Position next floor; keeping the lie drops it.

\textbf{When the Floor Clock Fills:} Floor becomes slick with spilled secrets; mirror eats a die from a pool; reflections turn hostile. A spilled secret reveals Utaran social regulations on proper courtship that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Floor becomes slick with spilled secrets. Among the secrets are references to Utaran noble families who lost their social standing during the empire's collapse.}
    \sbitem{2}{A mirror reflection speaks forbidden truths. The reflection wears the clothes of a Utaran aristocrat who died when the empire's social hierarchy collapsed.}
    \sbitem{3}{Mirror eats a die from a pool. The consumed die bears the markings of Utaran gaming traditions that were abandoned during wartime.}
    \sbitem{4+}{All reflections turn hostile, attacking with intimate knowledge. The reflections bear the faces of Utaran citizens who died during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when secrets spill or reflections turn hostile---omens of broken hearts, debts of emotional manipulation, echoes of Utaran social collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{VALERIAN DUSKHEART, STOLEN LOVE}
\nameimportance{Restoring his full identity acknowledges he was more than a trophy to be won. His family name connects to Utaran nobility who lost everything when the empire fell.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Backmask: Name spoken backward by mirror-choir; correct with counter-chant, the choir singing in the style of Utaran ceremonial traditions
    \item Palimpsest: His name visible only when true tears fall on his portrait, revealed beneath erased Utaran social invitations
    \item Emotional Key: Appears when a PC chooses honesty over victory, echoing the Utaran values of personal honor
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Portrait initials spell his name, with the initials arranged in Utaran noble family crest patterns
    \item[Actionable Lead:] Sway reveals his true feelings in hidden correspondence, including references to Utaran social gatherings that were cancelled during wartime
    \item[Aha! Proof:] Arcana work breaks the glamour to show his authentic self, revealing his true identity as a Utaran noble who became a refugee
\end{description}

\textbf{Extraction Test:} Sway or Arcana to see through the glamour (on fail, the illusion strengthens but reveals his middle name and family connections to Utaran refugee networks).

\textbf{Verification \& Use:} Speak his name while breaking the glamour focus; mispronunciations cause mirrors to shatter (harm 1) and add a clock segment. Speaking his full name correctly causes the ballroom to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Undo Glamour (Position +1)

\textbf{Restitution:} Keep the Lie (gain Favor now; lose a friend-later)

\textbf{Renunciation:} Break the Mirrors (Spade omen: shard sticks in soul)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Honest Heart (once: +1d to see through deception)}
\burden{Spade omen if you break mirrors; Ledger - Helped: Valerian if you undo glamour}
\textbf{Shard Kit:} Heart pendant grants +1d to recognize true love later; the pendant bears Utaran noble markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored his name; his family's refugee story connects to Utaran's fall
\textbf{Name Ledger Update:} Harmed: Valerian Duskheart • Helped: [ ] • Unresolved: [ ]

\entrytext{The ballroom's mirrors reflect not just your image, but accusations from every angle. Crystal chandeliers hang like frozen tears, and the air shimmers with the residue of broken enchantments. The ballroom's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran aristocratic venues that were shuttered during wartime emergencies.}

\exittext{The mirrors clear as Valerian's name is spoken. His reflection smiles sadly before fading, and the ballroom finds peace. The distant echo of the Noon Bell fades with satisfaction, and the venue remembers its former purpose as a place of honest fellowship.}
\end{floorbox}

% --- Floor 14 ---
\begin{floorbox}{14 --- The Broken Oath to the Dead}
\memoryline{He promised to guard a grave.}
\tonetags{contrition • hill • standing stones • broken promise}
\startposition{Dominant}
\floorclock{[4] --- Cold wind steals breath; name tries to climb into mouth; stones hum with power.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The promise had seemed simple---a favor for an old mentor, to watch over a grave where important research was buried. But as years passed and other opportunities arose, the vigil became inconvenient. He told himself the work was more important, that his research could benefit many more than honoring a single grave. But the truth was that convenience had trumped duty, and a sacred trust had been broken for personal gain. During the Siege of Ecktoria, such breaking of sacred trusts would become common as desperate officials prioritized immediate needs over long-term obligations.

\textbf{From the Harmed Party's Perspective:} The grave held more than just bones---it contained the culmination of a lifetime's work, knowledge that could benefit generations if properly protected and studied. When the guardian abandoned his post, when the sacred site was left vulnerable to grave-robbers and vandals, it felt like a betrayal of everything the deceased had worked for. Justice would have been honor---keeping the watch, protecting what had been entrusted. Many Utaran citizens faced similar betrayals as their empire's institutions failed to maintain their sacred obligations.

\textbf{Penance Sought:} To acknowledge his broken vigil and restore proper guardianship. The lich recognizes that his failure to keep a sacred trust mirrored the empire's failure to keep its promises to its citizens.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Hill with standing stones; an honored name is scratched out; disturbed earth shows signs of violation. The area's design reflects Utaran religious traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Standing stones bearing Utaran religious markings, scratched name with Utaran noble crests, disturbed grave with Utaran military honors, guardian's oath with Utaran legal seals, protective wards with Utaran magical symbols

\textbf{Senses:} The cold wind that steals breath; the hum of standing stones with lost power; the feel of violated sanctity; the distant echo of the Noon Bell that once marked time in Utaran religious sites.

\textbf{Adversaries/Agents:} Grave-robbers, ward spirits, the deceased's echo, stone guardians, the ghost of a Utaran religious administrator demanding proper respect for the dead.
\end{gmnote}

\handles{Survival (restore cairn, DV 3) • Rites (oath-binding, DV 2) • Skirmish (grave-robbers, DV 2) • Insight (why he failed, DV 3)}

\textbf{Position Drift:} Restoring the name raises Position next floor; sealing it away drops it.

\textbf{When the Floor Clock Fills:} Cold wind steals breath; name tries to climb into a PC's mouth; standing stones begin to crack. A cracked stone reveals Utaran religious regulations on proper grave guardianship that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Cold wind steals breath, causing Fatigue. The wind carries the essence of Utaran plague victims from the Siege of Ecktoria.}
    \sbitem{2}{Ward spirits grow agitated and hostile. The spirits bear the markings of Utaran religious traditions that were suppressed during wartime.}
    \sbitem{3}{A name tries to climb into a PC's mouth, demanding to be spoken. The name belongs to a Utaran official who died when the empire's religious institutions collapsed.}
    \sbitem{4+}{The deceased's echo rises, demanding explanation. The echo wears the robes of a Utaran religious figure who died during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the name tries to speak or when the deceased's echo rises---omens of broken vows, debts of violated trust, echoes of Utaran religious collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{MASTER ALDRIC THORNFIELD, GUARDIAN'S CHARGE}
\nameimportance{Restoring his full title acknowledges the respect and responsibility that was owed. His title connects to Utaran religious traditions, and his betrayal represents the empire's abandonment of its spiritual foundations.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of stone carvings spell his name, with carvings bearing Utaran religious symbols
    \item Palimpsest: His name visible only when proper offerings are made, revealed beneath erased Utaran religious texts
    \item Emotional Key: Appears when a PC chooses duty over convenience, echoing the Utaran values of honor and service
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Stone carvings arranged in his initials, with the arrangement matching Utaran religious honorifics
    \item[Actionable Lead:] Rites reveal his full title in burial records bearing Utaran religious seals
    \item[Aha! Proof:] Insight uncovers his true story in mentor's journals, including references to Utaran religious ceremonies that were cancelled during wartime
\end{description}

\textbf{Extraction Test:} Rites or Insight to restore his honor (on fail, the grave becomes more disturbed but reveals his middle name and religious credentials).

\textbf{Verification \& Use:} Speak his full name while restoring the cairn; mispronunciations cause the stones to hum painfully (harm 1) and add a clock segment. Speaking his full title correctly causes the hill to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Restore Name (gain Ancestor's Watch boon)

\textbf{Restitution:} Seal It Away (safer path opens; lose that boon)

\textbf{Renunciation:} Let the Dead Speak (learn truth; draw Deck now)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Ancestor's Watch (once: +1d to guardian duties)}
\burden{Loss of boon if you seal it away; Ledger - Helped: Master Aldric if you restore name}
\textbf{Shard Kit:} Guardian's oath fragment grants +1d to recognize duty later; the fragment bears Utaran religious markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored his name; his betrayal represents Utaran's religious collapse
\textbf{Name Ledger Update:} Harmed: Master Aldric Thornfield • Helped: [ ] • Unresolved: [ ]

\entrytext{The hill rises before you, crowned with standing stones that hum with residual power. But one name has been scratched out from the central stone, and the earth around the grave shows signs of recent disturbance. The area's construction features triple-course river-stone and laurel-circle motifs, reminiscent of Utaran religious sites that were desecrated during wartime conflicts.}

\exittext{The standing stones ring clearly as Master Aldric's name is spoken. The scratched-out letters fill in with golden light, and the hill finds peace. The distant echo of the Noon Bell fades with satisfaction, and the site remembers its former purpose as a place of sacred trust.}
\end{floorbox}

% --- Floor 15 ---
\begin{floorbox}{15 --- The Research Fire}
\memoryline{He set a fire to hide evidence.}
\tonetags{contrition • stacks • blaze • soot}
\startposition{Dominant}
\floorclock{[6] --- Roof beam falls; "saved" book is false; soot-spirits chant accusations.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The research had gone wrong---terribly, catastrophically wrong. Experiments had produced results he couldn't control, knowledge he couldn't contain. When authorities began asking questions, when colleagues started looking too closely at his work, he panicked. The fire seemed like a clean solution---destroy everything, start over with a new identity. But some secrets couldn't be burned away, and the truth had a way of rising from the ashes. During the Siege of Ecktoria, such desperate attempts to hide failures would become common as officials sought to conceal their mistakes from wartime oversight.

\textbf{From the Harmed Party's Perspective:} The library contained irreplaceable knowledge---centuries of accumulated wisdom, unique texts that existed nowhere else in the world. When it burned, when precious manuscripts turned to ash and knowledge was lost forever, it felt like a crime against humanity itself. Justice would have been accountability---facing the consequences, finding a way to make amends rather than destroying everything. Many Utaran institutions faced similar losses as their empire's knowledge repositories were destroyed during wartime conflicts.

\textbf{Penance Sought:} To acknowledge his attempt to erase the past and face the truth of his actions. The lich recognizes that his destruction of knowledge mirrored the empire's destruction of its own intellectual heritage.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Stacks blaze with magical fire; soot-spirits chant; precious books turn to ash while others reveal hidden truths. The library's design reflects Utaran scholarly traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Burning stacks with Utaran library classifications, soot-spirits bearing Utaran academic markings, fire focus with Utaran magical symbols, hidden safe-room with Utaran security features, ash-covered texts with Utaran scholarly seals

\textbf{Senses:} The smell of burning parchment and magic; the sound of soot-spirits chanting; the heat of destructive flames; the distant echo of the Noon Bell that once marked time in Utaran libraries.

\textbf{Adversaries/Agents:} Ash wraiths, fire elementals, soot-spirits, truth-seeking guardians, the ghost of a Utaran librarian demanding proper preservation of knowledge.
\end{gmnote}

\handles{Move (create firebreak, DV 3) • Tinker (sprinkler glyphs, DV 2) • Skirmish (ash wraiths, DV 2) • Insight (find the real truth he buried, DV 3)}

\textbf{Position Drift:} Saving lives raises Position next floor; saving proof drops it.

\textbf{When the Floor Clock Fills:} Roof beam falls; "saved" book is false; soot-spirits chant the names of all he tried to destroy. A fallen beam reveals Utaran library preservation protocols that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Magical flames spread unpredictably. The flames carry the essence of Utaran magical experiments that went wrong during the Siege of Ecktoria.}
    \sbitem{2}{Roof beam falls, blocking an exit. The beam bears the markings of Utaran construction standards that were compromised during wartime.}
    \sbitem{3}{A "saved" book is revealed to be false. The book contains forged Utaran historical records that were created to conceal wartime failures.}
    \sbitem{4+}{Soot-spirits chant the full list of destroyed knowledge. The list includes Utaran research that might have helped during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when beams fall or when soot-spirits chant---omens of hidden truths, debts of destroyed knowledge, echoes of Utaran intellectual collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE TRUTH OF THE NECROMANTIC EXPERIMENTS, BURIED IN ASHES}
\nameimportance{The truth itself deserved to be named and acknowledged, not destroyed. The truth connects to Utaran scholarly traditions, and its destruction represents the empire's abandonment of its intellectual heritage.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Palimpsest: Truth visible only when water is poured on ash-covered texts, revealed beneath erased Utaran academic publications
    \item Syllable Scavenger: Fragments of the truth found in partially burned documents bearing Utaran library classification systems
    \item Emotional Key: Appears when a PC chooses honesty over self-preservation, echoing the Utaran values of scholarly integrity
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Partially burned text spells "truth" in ash, with the text bearing Utaran academic terminology
    \item[Actionable Lead:] Insight reveals the scope of destroyed research, including references to Utaran defensive magic that might have helped during the Siege of Ecktoria
    \item[Aha! Proof:] Tinker work uncovers the hidden safe-room with complete records, including Utaran research that was suppressed during wartime
\end{description}

\textbf{Extraction Test:} Insight or Tinker to recover the truth (on fail, more knowledge is lost but reveals research topics and connections to Utaran intellectual collapse).

\textbf{Verification \& Use:} Speak the truth's name while preserving what remains; mispronunciations cause the fire to spread (harm 1) and add a clock segment. Speaking the full name correctly causes the library to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Save Lives (Blessing)

\textbf{Restitution:} Save Proof (unlock clue elsewhere; mark Guilt)

\textbf{Renunciation:} Let it Burn (Position -1; but gain Power cipher later)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Preservation Blessing (once: +1d to save valuable items)}
\burden{Guilt marker if you save proof; Ledger - Helped: Knowledge if you save lives}
\textbf{Shard Kit:} Ash-covered text grants +1d to recognize hidden truths later; the text bears Utaran scholarly seals
\textbf{Links:} Reduces Final Persuasion DV by 1 if you preserved knowledge; the destruction represents Utaran's intellectual collapse
\textbf{Name Ledger Update:} Harmed: Lost knowledge of the great library • Helped: [ ] • Unresolved: [ ]

\entrytext{Flames dance between the book stacks, fed by more than mere wood and paper. Soot-spirits chant in voices like rustling pages, and the air crackles with the destruction of irreplaceable knowledge. The library's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran academic institutions that were destroyed during wartime conflicts.}

\exittext{The flames dim as the truth's name is spoken. Ash settles into peaceful patterns, and among the ruins, new growth begins to emerge. The distant echo of the Noon Bell fades with satisfaction, and the library remembers its former purpose as a place of learning.}
\end{floorbox}

% --- Floor 16 ---
\begin{floorbox}{16 --- The Orphan's Blade}
\memoryline{He trained a child assassin.}
\tonetags{contrition • rooftop • bell-lanterns • corruption}
\startposition{Dominant}
\floorclock{[5] --- Knife flashes at throat; kill-phrase echoes; bell-lanterns swing wildly.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The child was a blank slate, easily molded, desperate for purpose and belonging. He told himself he was offering salvation---a chance to matter, to be more than a street urchin scraping for survival. But the truth was darker: he needed a tool, someone he could control completely, whose loyalty was absolute. In creating the perfect assassin, he had destroyed a child's chance at innocence and humanity. During the Siege of Ecktoria, such exploitation of the vulnerable would become common as desperate officials sought any advantage.

\textbf{From the Harmed Party's Perspective:} Before the training, there had been dreams---simple, childish hopes of growing up to be a hero or a scholar or anything but what they became. When those dreams were systematically stripped away, when violence was taught as love and obedience as virtue, something fundamental was lost. Justice would have been childhood---the right to grow up whole, to choose one's own path without manipulation. Many Utaran children faced similar corruptions as their empire's social services collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his corruption of innocence and break the chains of control he imposed. The lich recognizes that his exploitation of a child mirrored the empire's exploitation of its most vulnerable citizens.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Rooftop training ground; bell-lanterns swing in the wind; practice dummies bear the scars of countless lessons. The area's design reflects Utaran military traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Training dummies bearing Utaran military markings, bell-lanterns with Utaran noble crests, kill-phrase focus with Utaran magical symbols, childhood mementos with Utaran cultural markings, assassin's blade with Utaran weapon smithing marks

\textbf{Senses:} The clang of practice weapons; the whisper of wind through lanterns; the metallic taste of fear; the distant echo of the Noon Bell that once marked time in Utaran military training grounds.

\textbf{Adversaries/Agents:} Orphan assassins, bell-lantern spirits, kill-phrase echoes, childhood memories made manifest, the ghost of a Utaran military trainer demanding proper conduct.
\end{gmnote}

\handles{Sway (deprogram, DV 3) • Skirmish (disarm gently, DV 2) • Acrobatics (catch the fall, DV 3) • Arcana (unbind kill-phrase, DV 2)}

\textbf{Position Drift:} Breaking the phrase raises Position next floor; finishing the lesson drops it.

\textbf{When the Floor Clock Fills:} Knife flashes at a throat; kill-phrase echoes in the bells; bell-lanterns swing wildly, blinding with their light. A practice dummy reveals Utaran military training manuals that were used during the Siege of Ecktoria.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{A practice dummy animates, attacking with mechanical precision. The dummy bears the markings of Utaran military equipment that was requisitioned during wartime.}
    \sbitem{2}{Knife flashes at a throat, demanding a choice. The knife bears the markings of Utaran weapon smiths who died during the empire's collapse.}
    \sbitem{3}{Kill-phrase echoes in the bells, compelling obedience. The phrase echoes the commands used in Utaran military training that was intensified during wartime.}
    \sbitem{4+}{The child's spirit appears, demanding release from their conditioning. The spirit wears the clothes of a Utaran orphan who died during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the kill-phrase activates or when the child's spirit appears---omens of corrupted innocence, debts of stolen childhood, echoes of Utaran social collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE NAMELESS ONE, CHILD OF SHADOW}
\nameimportance{Restoring their identity acknowledges they were more than just a weapon. Their name connects to Utaran social traditions, and their corruption represents the empire's abandonment of its children.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Syllable Scavenger: Fragments of their true name found in childhood mementos bearing Utaran cultural markings
    \item Palimpsest: Their name visible only when tears fall on the assassin's blade, revealed beneath erased Utaran military records
    \item Emotional Key: Appears when a PC chooses compassion over control, echoing the Utaran values of protecting the innocent
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Childhood drawings spell out "Hope", with the drawings bearing Utaran artistic styles
    \item[Actionable Lead:] Sway reveals their true feelings in hidden letters, including references to Utaran social services that were cut during wartime
    \item[Aha! Proof:] Arcana work breaks the conditioning to show their authentic self, revealing their true identity as a Utaran orphan who became a refugee
\end{description}

\textbf{Extraction Test:} Sway or Arcana to see through the conditioning (on fail, the conditioning strengthens but reveals their original name and family connections to Utaran refugee networks).

\textbf{Verification \& Use:} Speak their true name while breaking the kill-phrase focus; mispronunciations cause the bell-lanterns to attack (harm 1) and add a clock segment. Speaking their full name correctly causes the rooftop to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Break the Phrase (gain Guarded Word talisman)

\textbf{Restitution:} Finish the Lesson (gain deadly ally; mark Stain)

\textbf{Renunciation:} Adopt Responsibility (downtime obligation reduces future DV once)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Guarded Word Talisman (once: immune to compulsion phrases)}
\burden{Stain if you finish the lesson; Ledger - Helped: Nameless One if you break the phrase}
\textbf{Shard Kit:} Childhood memento grants +1d to recognize manipulation later; the memento bears Utaran cultural markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their name; their corruption represents Utaran's social collapse
\textbf{Name Ledger Update:} Harmed: Nameless One • Helped: [ ] • Unresolved: [ ]

\entrytext{The rooftop training ground stretches before you, marked by countless scars where blade met blade. Bell-lanterns swing in the wind, casting dancing shadows that seem almost alive with remembered violence. The area's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran military training facilities that were converted for wartime use.}

\exittext{The bell-lanterns still as the child's true name is spoken. Their blade falls harmlessly to the ground, and for a moment, you see a child's smile instead of an assassin's glare. The distant echo of the Noon Bell fades with satisfaction, and the training ground remembers its former purpose as a place of proper instruction.}
\end{floorbox}

% --- Floor 17 ---
\begin{floorbox}{17 --- The Siege Engine Test}
\memoryline{He field-tested a weapon on a village.}
\tonetags{contrition • ballista field • stones • screams}
\startposition{Dominant}
\floorclock{[6] --- Misfire blasts the party; sponsor-echo offers coin; villagers' ghosts rise.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The weapon was revolutionary---capable of leveling fortifications, ending sieges before they began. He told himself the test was necessary, that the few lives lost were justified by the thousands that would be saved in future conflicts. But the villagers had been real people with real families, not abstract targets or acceptable losses. The screams he had tried to forget still haunted his dreams. During the Siege of Ecktoria, such testing of weapons on civilians would become common as desperate officials sought any advantage.

\textbf{From the Harmed Party's Perspective:} They had been going about their daily lives---tending fields, caring for children, preparing meals---when the sky suddenly erupted with destruction. The weapon came without warning, without declaration of war, without any recognition of their humanity. Justice would have been choice---the right to flee, to surrender, to exist without being collateral in someone else's ambitions. Many Utaran citizens faced similar destructions as their empire's weapons were tested on their own people during wartime.

\textbf{Penance Sought:} To acknowledge his role in their deaths and take responsibility for the weapon he created. The lich recognizes that his creation of destructive weapons mirrored the empire's creation of instruments of its own destruction.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Abandoned ballista field; stones weep with residual sorrow; broken homes dot the landscape like broken teeth. The area's design reflects Utaran military engineering---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Siege engine bearing Utaran military markings, weeping stones with Utaran religious symbols, victim memorials with Utaran civic seals, weapon schematics with Utaran engineering stamps, sponsor's contract with Utaran legal seals

\textbf{Senses:} The metallic taste of blood and fear; the sound of phantom screams; the feel of destructive power; the distant echo of the Noon Bell that once marked time in Utaran military installations.

\textbf{Adversaries/Agents:} Stone weepers (animated rocks that cry tears of acid), village ghosts, weapon engineers, sponsor's enforcer, the ghost of a Utaran military inspector demanding proper testing protocols.
\end{gmnote}

\handles{Tinker (sabotage, DV 3) • Command (evacuate echoes, DV 2) • Skirmish (fight engineers, DV 3) • Evidence (expose sponsor, DV 2)}

\textbf{Position Drift:} Sabotaging the weapon raises Position next floor; taking the coin drops it.

\textbf{When the Floor Clock Fills:} Misfire blasts the party; sponsor-echo offers coin to walk away; villagers' ghosts rise to demand justice. A broken home reveals Utaran civilian protection regulations that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{A stone weeper's tears cause acid burns. The tears carry the essence of Utaran plague victims from the Siege of Ecktoria.}
    \sbitem{2}{Misfire blasts the party, causing collateral damage. The misfire echoes the accidental destruction that occurred during Utaran weapons testing during wartime.}
    \sbitem{3}{A sponsor-echo offers coin to walk away from the truth. The sponsor bears the markings of Utaran military contractors who profited during wartime.}
    \sbitem{4+}{The villagers' ghosts rise, demanding justice for their deaths. The ghosts wear the clothes of Utaran civilians who died when their village was used for weapons testing during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the weapon misfires or when ghosts rise---omens of destructive ambition, debts of innocent blood, echoes of Utaran military collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE VILLAGE OF THORNWICK, DESTROYED BY PROGRESS}
\nameimportance{The village had an identity beyond being a test site. Its name connects to Utaran rural traditions, and its destruction represents the empire's betrayal of its own citizens.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of memorial stones spell the village name, with stones bearing Utaran municipal markings
    \item Palimpsest: Village name visible only when proper respects are paid to memorials, revealed beneath erased Utaran property records
    \item Emotional Key: Appears when a PC chooses accountability over advancement, echoing the Utaran values of protecting the innocent
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Memorial stones arranged in village initials, with the arrangement matching Utaran municipal signage
    \item[Actionable Lead:] Evidence reveals the full extent of the atrocity in military reports bearing Utaran defense ministry seals
    \item[Aha! Proof:] Command work uncovers the village's true history from survivor accounts, including references to Utaran civilian protection measures that were suspended during wartime
\end{description}

\textbf{Extraction Test:} Evidence or Command to expose the full truth (on fail, more victims are forgotten but reveal village details and connections to Utaran wartime policies).

\textbf{Verification \& Use:} Speak the village's name while sabotaging the weapon; mispronunciations cause the siege engine to target allies (harm 1) and add a clock segment. Speaking the full name correctly causes the field to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Sabotage (lower later harm)

\textbf{Restitution:} Expose (gain Leverage; attract enemies)

\textbf{Renunciation:} Take the Coin (resources now; Position -1 next two floors)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Weapon Saboteur (once: +1d to disable destructive devices)}
\burden{Position -1 if you take the coin; Ledger - Helped: Thornwick if you sabotage}
\textbf{Shard Kit:} Memorial stone fragment grants +1d to recognize war crimes later; the fragment bears Utaran civic seals
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their memory; the village's destruction represents Utaran's military collapse
\textbf{Name Ledger Update:} Harmed: Village of Thornwick • Helped: [ ] • Unresolved: [ ]

\entrytext{The abandoned ballista field stretches before you, its massive weapon pointing toward broken homes and weeping stones. The air hums with the residual power of destruction, and phantom screams echo in the wind. The area's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran military installations that were converted for wartime use.}

\exittext{The siege engine falls silent as Thornwick's name is spoken. The weeping stones dry their tears, and the broken village finds peace. The distant echo of the Noon Bell fades with satisfaction, and the field remembers its former purpose as a place of proper testing.}
\end{floorbox}

% --- Floor 18 ---
\begin{floorbox}{18 --- The University Vote}
\memoryline{He blocked reparations to a harmed cohort.}
\tonetags{contrition • senate • candles • ballots}
\startposition{Dominant}
\floorclock{[4] --- Whip flips votes; clock gains segment; academic censors arrive.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The reparations would have set a dangerous precedent---acknowledging fault, paying for mistakes, opening the floodgates for every disgruntled student or colleague to demand compensation. He told himself he was protecting the institution's integrity, its reputation, its future. But the truth was that pride and fear of financial loss had driven him to silence those who had been genuinely harmed by his colleagues' actions. During the Siege of Ecktoria, such blocking of accountability would become common as desperate officials sought to maintain institutional stability.

\textbf{From the Harmed Party's Perspective:} They had trusted the university to be a place of learning, growth, and safety. When that trust was betrayed, when they were harmed by those sworn to protect them, they sought justice through proper channels. When those channels were blocked by political maneuvering and bureaucratic indifference, it felt like a second betrayal. Justice would have been acknowledgment---recognizing the harm done and taking concrete steps to make amends. Many Utaran institutions faced similar failures as their empire's accountability mechanisms collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his role in silencing victims and support their quest for justice. The lich recognizes that his blocking of accountability mirrored the empire's blocking of its own citizens' cries for justice.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Senate in candle-gloom; ballots bleed wax; academic robes hang like accusations from empty chairs. The chamber's design reflects Utaran academic traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Voting chamber with Utaran university seals, bleeding ballots bearing Utaran administrative markings, victim testimonies with Utaran legal stamps, university charter with Utaran educational seals, academic robes with Utaran honorific markings

\textbf{Senses:} The scent of old wax and fear; the sound of whispered votes; the feel of institutional power; the distant echo of the Noon Bell that once marked time in Utaran academic institutions.

\textbf{Adversaries/Agents:} Academic censors, voting manipulators, victim echoes, university bureaucrats, the ghost of a Utaran educational administrator demanding proper procedure.
\end{gmnote}

\handles{Sway (coalition-build, DV 3) • Deception (parliamentary trick, DV 2) • Law (craft binding motion, DV 3) • Skirmish (duel a bully-doctor, DV 2)}

\textbf{Position Drift:} Passing restitution raises Position next floor; tabling it drops it.

\textbf{When the Floor Clock Fills:} Whip flips votes; clock gains a segment as rules shift; academic censors arrive to enforce silence. A ballot reveals Utaran academic regulations on proper grievance procedures that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{A voting whip flips votes unexpectedly. The whip bears the markings of Utaran administrative tools that were requisitioned during wartime.}
    \sbitem{2}{Academic censors arrive, demanding compliance. The censors wear the uniforms of Utaran educational officials who enforced wartime restrictions.}
    \sbitem{3}{Clock gains a segment as parliamentary rules shift against you. The rules echo the emergency measures implemented in Utaran institutions during the Siege of Ecktoria.}
    \sbitem{4+}{The victims' collective spirit rises, demanding justice. The spirit wears the robes of Utaran students who died when their university was converted for wartime use.}
\end{description}

\deckdraw{Draw when votes are manipulated or when victim spirits rise---omens of institutional betrayal, debts of silenced voices, echoes of Utaran educational collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE HARMED STUDENTS OF THE ARCANE PRACTICUM, SILENCED}
\nameimportance{They were individuals with specific grievances, not an abstract problem to be managed. Their collective name connects to Utaran academic traditions, and their silencing represents the empire's abandonment of its educational values.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of student ID numbers spell their collective identity, with numbers bearing Utaran university identification systems
    \item Palimpsest: Their names visible only when proper testimonies are read aloud, revealed beneath erased Utaran academic records
    \item Emotional Key: Appears when a PC chooses advocacy over institutional loyalty, echoing the Utaran values of protecting the vulnerable
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Student ID numbers arranged in chronological order, with the arrangement matching Utaran academic enrollment systems
    \item[Actionable Lead:] Law reveals their full case in university records bearing Utaran educational seals
    \item[Aha! Proof:] Sway work uncovers their individual stories from hidden testimonies, including references to Utaran student protection measures that were suspended during wartime
\end{description}

\textbf{Extraction Test:} Sway or Law to build support for their cause (on fail, more opposition arises but reveals student names and connections to Utaran educational collapse).

\textbf{Verification \& Use:} Speak their collective name while presenting their case; mispronunciations cause the academic censors to turn hostile (harm 1) and add a clock segment. Speaking their full name correctly causes the senate to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Pass Restitution (gain Public Grace)

\textbf{Restitution:} Table It (reduce heat; lose Grace)

\textbf{Renunciation:} Burn the Charter (scene shift now; future boon vs bureaucracy)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Academic Advocacy (once: +1d to support student causes)}
\burden{Loss of Grace if you table it; Ledger - Helped: Students if you pass restitution}
\textbf{Shard Kit:} University charter fragment grants +1d to recognize institutional corruption later; the fragment bears Utaran educational seals
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their cause; their silencing represents Utaran's educational collapse
\textbf{Name Ledger Update:} Harmed: Students of the Arcane Practicum • Helped: [ ] • Unresolved: [ ]

\entrytext{The senate chamber sits in candle-gloom, its empty chairs draped in academic robes like silent accusations. Ballots on the table bleed wax that spells out forgotten promises, and the air hums with the weight of unspoken truths. The chamber's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran university buildings that were converted for wartime administration.}

\exittext{The candles brighten as the students' names are spoken. The bleeding ballots seal themselves with golden wax, and the senate chamber finds justice. The distant echo of the Noon Bell fades with satisfaction, and the chamber remembers its former purpose as a place of proper governance.}
\end{floorbox}

% --- Floor 19 ---
\begin{floorbox}{19 --- The Expedition}
\memoryline{He misled peers into a lethal dig.}
\tonetags{contrition • necropolis • tent • sand}
\startposition{Dominant}
\floorclock{[6] --- Trench collapses; plague-sigil coughs; ancient guardian stirs.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The discovery promised to revolutionize their understanding of ancient magic, to bring fame and fortune to all involved. He had genuinely believed in the site's potential, had been convinced by fragmentary evidence that something extraordinary lay buried beneath the sands. But his certainty had been misplaced. The site was cursed, the warnings were real, and his ambition had led to the deaths of everyone he had brought along. During the Siege of Ecktoria, such reckless pursuits of knowledge would become common as desperate scholars sought breakthroughs at any cost.

\textbf{From the Harmed Party's Perspective:} They had trusted his expertise, believed in his research, and followed him into the unknown based on his assurances of safety and significance. When the tomb proved to be a death trap, when ancient wards they didn't understand activated and claimed their lives, it felt like a betrayal of that trust. Justice would have been caution---proper investigation, respect for local warnings, understanding that some knowledge comes with deadly prices. Many Utaran archaeological teams faced similar tragedies as their empire's desperate need for magical advantages led them into increasingly dangerous territories.

\textbf{Penance Sought:} To acknowledge his role in their deaths and ensure no others would suffer the same fate. The lich recognizes that his pursuit of glory at any cost mirrored the empire's own reckless pursuit of power that ultimately led to its downfall.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Wind-scoured necropolis; half-excavated tombs; expedition tents flap in unnatural wind. The area's design reflects Utaran archaeological traditions---triple-course river-stone foundations with laurel-circle motifs and Utaran civic symbols.

\textbf{Anchors:} Excavation tools bearing Utaran university markings, cursed artifacts with Utaran magical symbols, team roster with Utaran academic stamps, expedition log with Utaran scholarly seals, sand-ghouls with Utaran necromantic markings

\textbf{Senses:} The taste of ancient dust and copper; the sound of shifting sand like whispers; the feel of violated tombs; the distant echo of the Noon Bell that once marked time in Utaran archaeological sites.

\textbf{Adversaries/Agents:} Sand-ghouls, cursed artifacts, plague spirits, tomb guardians, the ghost of a Utaran archaeological administrator demanding proper excavation protocols.
\end{gmnote}

\handles{Arcana (identify curse, DV 3) • Survival (navigate tomb, DV 2) • Skirmish (fight guardians, DV 3) • Insight (understand true nature, DV 2)}

\textbf{Position Drift:} Proper respect raises Position next floor; plundering drops it.

\textbf{When the Floor Clock Fills:} Trench collapses; plague-sigil coughs, spreading disease; ancient guardian stirs from its slumber. A collapsed trench reveals Utaran archaeological safety protocols that were abandoned during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Sand shifts unpredictably, creating hazards. The shifting echoes the unstable political situation that plagued Utaran archaeological expeditions during the Siege of Ecktoria.}
    \sbitem{2}{A trench collapses, trapping party members. The collapse echoes the cave-ins that killed Utaran archaeological teams during the Siege of Ecktoria.}
    \sbitem{3}{A plague-sigil coughs, spreading disease. The sigil bears the markings of Utaran magical experiments that went wrong during wartime.}
    \sbitem{4+}{The expedition's ghosts rise, demanding accountability for their deaths. The ghosts wear the uniforms of Utaran scholars who died when their research expedition was attacked during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the trench collapses or when ghosts rise---omens of archaeological hubris, debts of betrayed trust, echoes of Utaran scholarly collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE EXPEDITION TO THE TOMB OF THE FORGOTTEN PHARAOHS, BETRAYED}
\nameimportance{The expedition had a purpose and identity beyond being a source of cursed artifacts. Its name connects to Utaran archaeological traditions, and its betrayal represents the empire's abandonment of its scholarly values.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of team member names spell the expedition's true purpose, with names bearing Utaran university identification systems
    \item Palimpsest: Expedition name visible only when proper funeral rites are performed, revealed beneath erased Utaran archaeological records
    \item Emotional Key: Appears when a PC chooses honesty over glory, echoing the Utaran values of scholarly integrity
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Team roster arranged in alphabetical order, with the arrangement matching Utaran academic enrollment systems
    \item[Actionable Lead:] Arcana reveals the true nature of the site in ancient records bearing Utaran library stamps
    \item[Aha! Proof:] Survival work uncovers the expedition's original noble intentions, including references to Utaran archaeological protocols that were suspended during wartime
\end{description}

\textbf{Extraction Test:} Arcana or Survival to understand what went wrong (on fail, more cursed artifacts are disturbed but reveal expedition details and connections to Utaran scholarly collapse).

\textbf{Verification \& Use:} Speak the expedition's name while performing proper memorial rites; mispronunciations cause the sand-ghouls to turn hostile (harm 1) and add a clock segment. Speaking the full name correctly causes the necropolis to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Seal the Tomb (Blessing vs miasma)

\textbf{Restitution:} Plunder (gain Relic; mark Curse)

\textbf{Renunciation:} Name the Dead (lower Persuasion DV later)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Tomb Guardian's Blessing (once: +1d vs undead in burial sites)}
\burden{Curse marker if you plunder; Ledger - Helped: Expedition if you seal tomb}
\textbf{Shard Kit:} Expedition log page grants +1d to recognize dangerous sites later; the page bears Utaran scholarly seals
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their memory; the expedition's betrayal represents Utaran's scholarly collapse
\textbf{Name Ledger Update:} Harmed: Expedition team members • Helped: [ ] • Unresolved: [ ]

\entrytext{The wind-scoured necropolis stretches before you, its half-excavated tombs yawning like broken mouths. Tent pegs creak with ancient malice, and the air tastes of disturbed earth and forgotten deaths. The area's construction features triple-course river-stone foundations and laurel-circle motifs, reminiscent of Utaran archaeological sites that were desecrated during wartime conflicts.}

\exittext{The sand settles as the expedition's name is spoken. The tombs close respectfully, and the necropolis finds peace. The distant echo of the Noon Bell fades with satisfaction, and the site remembers its former purpose as a place of proper study.}
\end{floorbox}

% --- Floor 20 ---
\begin{floorbox}{20 --- The Confessor's Lament}
\memoryline{He used a lover's confession in a paper.}
\tonetags{contrition • confessional • glass • ink}
\startposition{Dominant}
\floorclock{[4] --- Ink stains hands; echo asks for PC confession; privacy barrier weakens.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The confession had contained insights---raw, emotional truths that could advance psychological understanding, contribute to academic discourse. He told himself he was anonymizing the content, removing identifying details, serving the greater good of knowledge. But the truth was that intimacy had been weaponized, trust betrayed for personal advancement. The lover had shared their soul in confidence, and he had turned it into data points and publishable material. During the Siege of Ecktoria, such betrayals of trust would become common as desperate officials sought any advantage.

\textbf{From the Harmed Party's Perspective:} The confessional had been sacred space---a place where vulnerabilities could be shared without fear of judgment or exposure. When those private words were used without consent, when intimate moments became public property, it felt like a violation of the soul itself. Justice would have been respect---honoring the boundaries of trust, keeping confidences inviolate. Many Utaran citizens faced similar betrayals as their empire's security apparatus turned inward during wartime.

\textbf{Penance Sought:} To acknowledge his breach of confidence and restore the sanctity of private trust. The lich recognizes that his violation of privacy mirrored the empire's violation of its citizens' fundamental rights.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Confessional of glass; ink drips inside like tears; privacy barriers flicker between transparency and opacity. The confessional's design reflects Utaran religious traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Glass confessional bearing Utaran religious markings, dripping ink with Utaran scholarly symbols, published paper with Utaran academic stamps, lover's portrait with Utaran artistic styles, privacy wards with Utaran magical symbols

\textbf{Senses:} The metallic taste of betrayal; the sound of dripping ink like falling tears; the feel of exposed vulnerability; the distant echo of the Noon Bell that once marked time in Utaran religious confessionals.

\textbf{Adversaries/Agents:} Privacy wraiths (spirits that embody violated confidentiality), confession echoes, academic censors, ink elementals, the ghost of a Utaran religious confessor demanding proper respect for sacred trust.
\end{gmnote}

\handles{Presence (apology, DV 2) • Evidence (destroy the paper, DV 3) • Deception (convince the echo it was mutual, DV 3) • Arcana (blank the confession, DV 2)}

\textbf{Position Drift:} Public apology raises Position next floor; quiet erasure drops it.

\textbf{When the Floor Clock Fills:} Ink stains a PC's hands (Stain); echo asks a PC to confess something true; privacy barrier weakens completely. A stained paper reveals Utaran regulations on academic ethics that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Ink stains a PC's hands, marking them with Stain. The stain carries the essence of Utaran scholars who were denounced for maintaining ethical standards during the Siege of Ecktoria.}
    \sbitem{2}{Privacy barrier flickers, exposing private thoughts. The exposure echoes the surveillance that overtook Utaran society during wartime.}
    \sbitem{3}{The echo asks a PC to confess something true, putting them on the spot. The demand echoes the forced confessions used by Utaran security forces during wartime.}
    \sbitem{4+}{Privacy wraiths manifest, demanding justice for violated confidences. The wraiths bear the faces of Utaran citizens who died in security prisons during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when hands are stained or when privacy wraiths appear---omens of violated trust, debts of exposed intimacy, echoes of Utaran social collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{ELARA MORNINGSTAR, VIOLATED IN CONFIDENCE}
\nameimportance{Restoring her full identity acknowledges she was more than just a source of data. Her name connects to Utaran noble traditions, and her violation represents the empire's betrayal of its citizens' fundamental rights.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Palimpsest: Her name visible only when tears fall on the published paper, revealed beneath erased Utaran social invitations
    \item Emotional Key: Appears when a PC chooses honesty over academic advancement, echoing the Utaran values of personal honor
    \item Syllable Scavenger: Fragments of her name found in redacted portions of the paper bearing Utaran academic classification systems
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Redacted portions spell her initials when held to light, with the redactions matching Utaran academic censorship patterns
    \item[Actionable Lead:] Presence reveals her true feelings in hidden correspondence, including references to Utaran social gatherings that were cancelled during wartime
    \item[Aha! Proof:] Arcana work uncovers her full identity in the paper's metadata, including her family connections to Utaran refugee networks
\end{description}

\textbf{Extraction Test:} Presence or Arcana to restore her dignity (on fail, more private information is exposed but reveals her middle name and family connections to Utaran noble houses).

\textbf{Verification \& Use:} Speak her full name while destroying the paper; mispronunciations cause the ink to attack (harm 1) and add a clock segment. Speaking her full name correctly causes the confessional to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Public Apology (gain Truth's Edge die once)

\textbf{Restitution:} Quiet Erasure (safer now; debt later)

\textbf{Renunciation:} Trade Truths (everyone tells one; Boons +1)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Truth's Edge Die (once: add to any social pool)}
\burden{Future debt if you erase quietly; Ledger - Helped: Elara if you apologize publicly}
\textbf{Shard Kit:} Lover's portrait grants +1d to recognize manipulation later; the portrait bears Utaran noble markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her violation represents Utaran's social collapse
\textbf{Name Ledger Update:} Harmed: Elara Morningstar • Helped: [ ] • Unresolved: [ ]

\entrytext{The glass confessional stands before you, its transparent walls offering no privacy. Ink drips inside like tears of shame, and the air hums with the weight of violated confidences. The confessional's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran religious venues that were converted for wartime surveillance.}

\exittext{The ink stops dripping as Elara's name is spoken. The glass walls fog over respectfully, and the confessional finds privacy in memory. The distant echo of the Noon Bell fades with satisfaction, and the venue remembers its former purpose as a place of sacred trust.}
\end{floorbox}

% --- Floor 21 ---
\begin{floorbox}{21 --- The Coven Gate}
\memoryline{He stole a ritual and sold it.}
\tonetags{contrition • standing stones • fog • betrayal}
\startposition{Dominant}
\floorclock{[5] --- Buyer names an ally; stone groans; threshold tears.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The ritual was magnificent---complex, powerful, the culmination of decades of coven work. He told himself he was preserving knowledge that might otherwise be lost, that selling it to a responsible buyer would ensure its continued study. But the truth was simpler: he coveted what wasn't his, betrayed the trust of those who had shared their secrets, and profited from their collective wisdom like a common thief. During the Siege of Ecktoria, such theft of magical knowledge would become common as desperate officials sought any advantage.

\textbf{From the Harmed Party's Perspective:} The ritual was more than a spell---it was their heritage, their identity, the sacred work that bound them together as a community. When it was stolen and sold, when their carefully guarded secrets became commodities in the open market, it felt like a theft of their very souls. Justice would have been respect---honoring their ownership, seeking permission before sharing, understanding that some knowledge comes with strings attached. Many Utaran magical communities faced similar betrayals as their empire's magical resources were requisitioned for wartime use.

\textbf{Penance Sought:} To acknowledge his theft and restore what he took from the coven. The lich recognizes that his theft of magical knowledge mirrored the empire's theft of its citizens' cultural heritage.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Standing stones with borrowed sigils, a buyer waiting in fog; the air crackles with stolen power. The area's design reflects Utaran magical traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Standing stones bearing Utaran magical markings, borrowed sigils with Utaran coven symbols, buyer's contract with Utaran legal seals, coven grimoire with Utaran library stamps, threshold markers with Utaran religious symbols

\textbf{Senses:} The taste of stolen magic; the sound of groaning stone; the feel of violated sacred space; the distant echo of the Noon Bell that once marked time in Utaran magical sites.

\textbf{Adversaries/Agents:} Sigil thieves, fog wraiths, coven guardians, ritual buyers, threshold demons, the ghost of a Utaran magical administrator demanding proper respect for magical traditions.
\end{gmnote}

\handles{Arcana (restore rightful words, DV 3) • Sway (renegotiate, DV 2) • Skirmish (stop the exchange, DV 3) • Theft (swap the grimoire, DV 2)}

\textbf{Position Drift:} Returning it raises Position next floor; selling again drops it.

\textbf{When the Floor Clock Fills:} Buyer names an ally among the party; stone groans with violated power; threshold tears open to darker realms. A borrowed sigil reveals Utaran magical regulations that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{A sigil thief attempts to steal from the party. The thief bears the markings of Utaran magical researchers who were forced to steal during the Siege of Ecktoria.}
    \sbitem{2}{Buyer names an ally, creating internal conflict. The ally wears the uniform of a Utaran magical official who profited from wartime requisitions.}
    \sbitem{3}{Stone groans, causing environmental hazards. The groaning echoes the destruction of Utaran magical sites during wartime conflicts.}
    \sbitem{4+}{Threshold tears open, unleashing something ancient and hungry. The entity bears the markings of Utaran magical experiments that went wrong during wartime.}
\end{description}

\deckdraw{Draw when the threshold tears or when allies are named---omens of magical theft, debts of violated traditions, echoes of Utaran magical collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE COVEN OF THE SILVER MOON, BETRAYED}
\nameimportance{The coven had an identity beyond being a source of magical knowledge. Its name connects to Utaran magical traditions, and its betrayal represents the empire's abandonment of its mystical heritage.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of sigil symbols spell the coven's name, with symbols bearing Utaran magical classification systems
    \item Palimpsest: Coven name visible only when proper moonlight falls on standing stones, revealed beneath erased Utaran magical texts
    \item Emotional Key: Appears when a PC chooses community over individual gain, echoing the Utaran values of collective responsibility
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Sigil symbols arranged in coven initials, with the arrangement matching Utaran magical honorifics
    \item[Actionable Lead:] Arcana reveals their full identity in ritual texts bearing Utaran library stamps
    \item[Aha! Proof:] Sway work uncovers their true story from coven elders, including references to Utaran magical traditions that were suppressed during wartime
\end{description}

\textbf{Extraction Test:} Arcana or Sway to restore their honor (on fail, more sigils are corrupted but reveal coven details and connections to Utaran magical collapse).

\textbf{Verification \& Use:} Speak the coven's name while returning the ritual; mispronunciations cause the stones to attack (harm 1) and add a clock segment. Speaking the full name correctly causes the standing stones to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Return It (blessing vs wards)

\textbf{Restitution:} Sell Again (gain Coin; draw Spade omen)

\textbf{Renunciation:} Bind the Gate (close a later hazard)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Coven's Blessing (once: +1d vs magical betrayal)}
\burden{Spade omen if you sell again; Ledger - Helped: Silver Moon Coven if you return it}
\textbf{Shard Kit:} Sigil fragment grants +1d to recognize stolen magic later; the fragment bears Utaran magical symbols
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their name; the coven's betrayal represents Utaran's magical collapse
\textbf{Name Ledger Update:} Harmed: Coven of the Silver Moon • Helped: [ ] • Unresolved: [ ]

\entrytext{The standing stones form a perfect circle, their surfaces etched with sigils that don't quite belong. In the center, a figure waits in fog, holding out coin for secrets that were never theirs to sell. The area's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran magical sites that were requisitioned during wartime emergencies.}

\exittext{The stones ring clearly as the coven's name is spoken. The borrowed sigils fade away, and the circle finds peace. The distant echo of the Noon Bell fades with satisfaction, and the site remembers its former purpose as a place of proper magical practice.}
\end{floorbox}

% --- Floor 22 ---
\begin{floorbox}{22 --- The City That Starved}
\memoryline{He speculated on grain during famine.}
\tonetags{contrition • market • scales • hunger}
\startposition{Dominant}
\floorclock{[6] --- Spoilage spreads; grain-storm blinds; desperate mob gathers.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The market forces were clear---supply and demand, the natural fluctuation of resources. He told himself he was merely taking advantage of opportunities, that others would have done the same, that his actions were simply business. But while he profited from scarcity, families starved in the streets below. The weight of their hunger became the weight of his guilt, measured not in coin but in lives. During the Siege of Ecktoria, such profiteering would become common as desperate merchants sought to maximize profits during wartime shortages.

\textbf{From the Harmed Party's Perspective:} They had watched their children grow thin, had traded away everything of value for a few more days of survival. When merchants hoarded grain while people died, when profit was valued over human life, it felt like a betrayal of the most basic social contract. Justice would have been compassion---sharing resources, prioritizing community over coin, recognizing that some things transcend market value. Many Utaran citizens faced similar starvation as their empire's food distribution systems collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his profiteering from suffering and make amends for the hunger he enabled. The lich recognizes that his exploitation of famine mirrored the empire's exploitation of its citizens' desperation.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Market of empty baskets; scales hang lopsided with the weight of inequality; the air reeks of desperation. The market's design reflects Utaran commercial traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Empty baskets bearing Utaran merchant markings, lopsided scales with Utaran commercial symbols, grain ledgers with Utaran administrative numbering, famine memorials with Utaran civic seals, merchant contracts with Utaran legal stamps

\textbf{Senses:} The smell of decay and desperation; the sound of weeping and gnashing teeth; the feel of gnawing hunger; the distant echo of the Noon Bell that once marked time in Utaran marketplaces.

\textbf{Adversaries/Agents:} Hunger wraiths, desperate mobs, grain elementals, famine spirits, corrupt merchants, the ghost of a Utaran commercial administrator demanding proper market conduct.
\end{gmnote}

\handles{Trade (unlock grain, DV 3) • Command (food line order, DV 2) • Skirmish (stop mob, DV 3) • Rites (invoke charity rite, DV 2)}

\textbf{Position Drift:} Releasing stores raises Position next floor; protecting warehouse drops it.

\textbf{When the Floor Clock Fills:} Spoilage spreads to remaining supplies; grain-storm blinds with chaff and dust; desperate mob gathers with pitchforks and rage. A spoiled grain sack reveals Utaran food distribution records from the Siege of Ecktoria.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Spoilage spreads, destroying more food. The spoilage echoes the food shortages that plagued Utaran cities during wartime.}
    \sbitem{2}{Grain-storm blinds with chaff and dust. The storm carries the essence of Utaran famine victims from the Siege of Ecktoria.}
    \sbitem{3}{Desperate mob gathers, demanding food. The mob wears the clothes of Utaran citizens who died of starvation during wartime shortages.}
    \sbitem{4+}{Hunger wraiths manifest, embodying the famine's suffering. The wraiths bear the faces of Utaran officials who were executed for hoarding food during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the mob gathers or when hunger wraiths appear---omens of economic cruelty, debts of enabled suffering, echoes of Utaran social collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE STARVING CITIZENS OF RAVENHOLT, BETRAYED BY GREED}
\nameimportance{They were individuals with specific suffering, not an abstract economic concept. Their collective name connects to Utaran urban traditions, and their betrayal represents the empire's abandonment of its citizens' basic needs.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of famine dates spell the city's name, with dates bearing Utaran municipal record systems
    \item Palimpsest: City name visible only when proper charity rites are performed, revealed beneath erased Utaran civic records
    \item Emotional Key: Appears when a PC chooses compassion over profit, echoing the Utaran values of social responsibility
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Famine dates arranged chronologically, with the arrangement matching Utaran municipal record systems
    \item[Actionable Lead:] Trade reveals their full story in city records bearing Utaran administrative seals
    \item[Aha! Proof:] Command work uncovers their individual struggles from survivor accounts, including references to Utaran food distribution systems that failed during wartime
\end{description}

\textbf{Extraction Test:} Trade or Command to understand their suffering (on fail, more starvation occurs but reveals city details and connections to Utaran social collapse).

\textbf{Verification \& Use:} Speak the city's name while performing charity rites; mispronunciations cause the hunger wraiths to turn hostile (harm 1) and add a clock segment. Speaking the full name correctly causes the market to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Release Stores (Public Grace)

\textbf{Restitution:} Protect the Warehouse (resources; mark Stain)

\textbf{Renunciation:} Name the Brokers (reduce faction heat later; draw Diamond omen now)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Charity's Blessing (once: +1d to feed the hungry)}
\burden{Stain if you protect warehouse; Ledger - Helped: Ravenholt citizens if you release stores}
\textbf{Shard Kit:} Empty basket grants +1d to recognize economic injustice later; the basket bears Utaran merchant markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their memory; the city's betrayal represents Utaran's social collapse
\textbf{Name Ledger Update:} Harmed: Citizens of Ravenholt • Helped: [ ] • Unresolved: [ ]

\entrytext{The market square lies empty except for baskets that once held hope. The scales hang lopsided, weighted down by the measure of suffering, and the air reeks of desperation and decay. The area's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran marketplaces that were requisitioned during wartime emergencies.}

\exittext{The scales balance as Ravenholt's name is spoken. The empty baskets fill with simple grain, and the market finds justice in shared abundance. The distant echo of the Noon Bell fades with satisfaction, and the marketplace remembers its former purpose as a place of fair trade.}
\end{floorbox}

% --- Floor 23 ---
\begin{floorbox}{23 --- The High Stair}
\memoryline{He let a rival fall to claim a seat.}
\tonetags{contrition • endless stair • rails • ambition}
\startposition{Dominant}
\floorclock{[4] --- Boot slips; third climber grabs ankle; rival's echo calls for justice.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The seat represented everything---power, influence, the ability to shape policy and direct resources. He told himself his rival was unworthy, that the organization needed his vision, his leadership, his superior intellect. But when the moment came and he could have extended a hand, he chose to step over instead. The rival's fall echoed in his dreams for years, a reminder that ambition had triumphed over basic humanity. During the Siege of Ecktoria, such betrayals of colleagues would become common as desperate officials sought to secure their positions.

\textbf{From the Harmed Party's Perspective:} They had worked alongside him for years, sharing ideas, building projects, earning their place through merit and dedication. When the crucial moment came, when they needed help to reach their shared goal, the betrayal cut deeper than any physical wound. Justice would have been sportsmanship---fair competition, mutual respect, the understanding that success was better shared than stolen. Many Utaran officials faced similar betrayals as their empire's political structure collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his betrayal and honor the merit of his fallen rival. The lich recognizes that his betrayal of a colleague mirrored the empire's betrayal of its own officials' trust.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Endless stair with two rails; one is greased, the other worn smooth by honest hands; the climb seems to go on forever. The stair's design reflects Utaran administrative traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Endless stair bearing Utaran administrative markings, greased rail with Utaran political symbols, rival's belongings with Utaran official seals, seat of power with Utaran governmental symbols, climbing equipment with Utaran engineering stamps

\textbf{Senses:} The metallic taste of betrayal; the sound of slipping footsteps; the feel of treacherous footing; the distant echo of the Noon Bell that once marked time in Utaran administrative buildings.

\textbf{Adversaries/Agents:} Rival echoes, stair guardians, falling hazards, ambition demons, climbing competitors, the ghost of a Utaran administrative official demanding proper conduct.
\end{gmnote}

\handles{Move (save-the-rival, DV 3) • Sway (admit the truth, DV 2) • Skirmish (cut the saboteur cord, DV 3) • Tinker (fix the rail, DV 2)}

\textbf{Position Drift:} Lifting them raises Position next floor; stepping over drops it.

\textbf{When the Floor Clock Fills:} Your boot slips on the greased rail; a third climber grabs your ankle, trying to pull you down; the rival's echo calls for justice from below. A greased rail section reveals Utaran political maneuvering records from the Siege of Ecktoria.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Your boot slips, causing a dangerous fall. The fall echoes the political falls that plagued Utaran officials during wartime.}
    \sbitem{2}{A hidden saboteur tries to cut climbing ropes. The saboteur wears the uniform of a Utaran security official who was tasked with eliminating political rivals during the Siege of Ecktoria.}
    \sbitem{3}{A third climber grabs your ankle, creating a dilemma. The climber wears the clothes of a Utaran citizen who died when political infighting escalated during wartime.}
    \sbitem{4+}{The rival's echo rises, demanding accountability for their fall. The echo wears the robes of a Utaran official who died when political betrayal escalated during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when you slip or when the rival's echo rises---omens of betrayed trust, debts of stolen opportunity, echoes of Utaran political collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{MASTER THADDEUS IRONWRIGHT, BETRAYED IN ASCENT}
\nameimportance{Restoring his full title acknowledges his worth beyond being a stepping stone. His title connects to Utaran administrative traditions, and his betrayal represents the empire's abandonment of its meritocratic values.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of stair carvings spell his name, with carvings bearing Utaran administrative honorifics
    \item Palimpsest: His name visible only when honest hands touch the good rail, revealed beneath erased Utaran official records
    \item Emotional Key: Appears when a PC chooses integrity over advancement, echoing the Utaran values of merit and honor
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Stair carvings arranged in his initials, with the arrangement matching Utaran administrative honorifics
    \item[Actionable Lead:] Sway reveals his true contributions in organizational records bearing Utaran governmental seals
    \item[Aha! Proof:] Tinker work uncovers his plans that were stolen, including references to Utaran administrative reforms that were abandoned during wartime
\end{description}

\textbf{Extraction Test:} Sway or Tinker to restore his honor (on fail, more sabotage occurs but reveals his middle name and administrative credentials).

\textbf{Verification \& Use:} Speak his full name while fixing the rail; mispronunciations cause the stair to become more treacherous (harm 1) and add a clock segment. Speaking his full title correctly causes the stair to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Lift Them (Position +1)

\textbf{Restitution:} Admit It (DV -1 later)

\textbf{Renunciation:} Step Over (gain Seat token; Desperate start next floor)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Integrity's Reward (once: +1d to honest competition)}
\burden{Desperate start if you step over; Ledger - Helped: Thaddeus if you lift him}
\textbf{Shard Kit:} Climbing rope grants +1d to recognize sabotage later; the rope bears Utaran engineering markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored his name; his betrayal represents Utaran's political collapse
\textbf{Name Ledger Update:} Harmed: Master Thaddeus Ironwright • Helped: [ ] • Unresolved: [ ]

\entrytext{The endless stair stretches upward, its two rails telling different stories---one smooth and worn by honest hands, the other greased with the residue of betrayal. Above, the seat of power gleams temptingly, while below, shadows gather like accusations. The stair's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran administrative buildings that were converted for wartime use.}

\exittext{The stair's treacherous rail smooths as Thaddeus's name is spoken. The greased section fades away, and the climb becomes honest effort once more. The distant echo of the Noon Bell fades with satisfaction, and the stairway remembers its former purpose as a path to proper governance.}
\end{floorbox}

% --- Floor 24 ---
\begin{floorbox}{24 --- The Dragon's Loan}
\memoryline{He took a predatory loan that doomed a village.}
\tonetags{contrition • coin-hall • drake • debt}
\startposition{Dominant}
\floorclock{[5] --- Interest compounds; draconic omen opens eye; village echoes plead.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The research required funding---rare components, specialized equipment, assistants who understood the work. The dragon's offer had seemed reasonable at the time, the terms fair for the resources needed. He told himself he could repay, that the village's temporary hardship would be worth the breakthrough he could achieve. But the dragon's arithmetic was cruel, its patience endless, and the debt grew like a cancer, consuming everything the village held dear. During the Siege of Ecktoria, such predatory lending would become common as desperate officials sought resources by any means.

\textbf{From the Harmed Party's Perspective:} They had trusted him, believed in his vision, invested their hopes in his promises of a better future. When the dragon's loan came due, when their homes and livelihoods were seized for debts they had never agreed to, it felt like a betrayal by the very person they had supported. Justice would have been protection---shielding them from exploitation, finding ethical funding, understanding that some prices were too high to pay. Many Utaran villages faced similar exploitations as their empire's financial systems collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his role in their exploitation and free them from the debt he incurred. The lich recognizes that his exploitation of the village mirrored the empire's exploitation of its citizens' desperation.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Coin-hall with a sleeping drake bound in ink; golden ledgers stack like towers; the air hums with the weight of unpayable debt. The hall's design reflects Utaran commercial traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Sleeping drake bearing Utaran magical markings, golden ledgers with Utaran financial symbols, debt contracts with Utaran legal seals, village memorial with Utaran civic stamps, payment altar with Utaran religious symbols

\textbf{Senses:} The metallic taste of coin and fear; the sound of counting gold; the feel of crushing debt; the distant echo of the Noon Bell that once marked time in Utaran financial institutions.

\textbf{Adversaries/Agents:} Debt collectors, ink wyrms, village echoes, compound interest spirits, draconic enforcers, the ghost of a Utaran financial administrator demanding proper lending practices.
\end{gmnote}

\handles{Trade (renegotiate, DV 3) • Arcana (smudge sigil, DV 2) • Skirmish (fight collectors, DV 3) • Evidence (prove coercion, DV 2)}

\textbf{Position Drift:} Refinancing to mercy raises Position next floor; paying in blood drops it.

\textbf{When the Floor Clock Fills:} Interest compounds: add a segment; draconic omen opens its eye, watching for payment; village echoes plead for release from debt. A debt contract reveals Utaran financial regulations that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Interest compounds, making debt worse. The compounding echoes the financial collapses that plagued Utaran cities during wartime.}
    \sbitem{2}{Compound interest spirits emerge, demanding payment. The spirits bear the faces of Utaran merchants who died when their businesses failed during wartime economic collapse.}
    \sbitem{3}{Draconic omen opens its eye, increasing pressure. The omen echoes the draconian measures implemented by Utaran authorities during wartime financial crises.}
    \sbitem{4+}{Village echoes rise, pleading for debt relief. The echoes wear the clothes of Utaran villagers who lost everything to predatory lending during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when interest compounds or when village echoes rise---omens of exploitative debt, debts of enabled oppression, echoes of Utaran economic collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE VILLAGE OF GOLDENFIELDS, ENSLAVED BY DEBT}
\nameimportance{The village had an identity beyond being collateral for a loan. Its name connects to Utaran rural traditions, and its enslavement represents the empire's abandonment of its citizens' economic freedom.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of crop names spell the village name, with names bearing Utaran agricultural classification systems
    \item Palimpsest: Village name visible only when proper payment is made, revealed beneath erased Utaran property records
    \item Emotional Key: Appears when a PC chooses community over individual advancement, echoing the Utaran values of economic justice
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Crop names arranged in village initials, with the arrangement matching Utaran agricultural record systems
    \item[Actionable Lead:] Trade reveals their full story in financial records bearing Utaran treasury seals
    \item[Aha! Proof:] Evidence uncovers their true situation in auditor reports, including references to Utaran agricultural support systems that failed during wartime
\end{description}

\textbf{Extraction Test:} Trade or Evidence to understand their plight (on fail, more debt accumulates but reveals village details and connections to Utaran economic collapse).

\textbf{Verification \& Use:} Speak the village's name while making proper payment; mispronunciations cause the drake to awaken (harm 1) and add a clock segment. Speaking the full name correctly causes the coin-hall to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Refinance to Mercy (lower future Diamond DVs)

\textbf{Restitution:} Pay in Blood (harm now; later grace)

\textbf{Renunciation:} Burn the Ledger (scene shift; draw Deck now)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Debt Freedom (once: immune to predatory lending)}
\burden{Harm if you pay in blood; Ledger - Helped: Goldenfields if you refinance mercifully}
\textbf{Shard Kit:} Golden ledger page grants +1d to recognize exploitative contracts later; the page bears Utaran financial markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their name; the village's enslavement represents Utaran's economic collapse
\textbf{Name Ledger Update:} Harmed: Village of Goldenfields • Helped: [ ] • Unresolved: [ ]

\entrytext{The coin-hall stretches before you, its golden ledgers stacked like towers of temptation. At the center, a drake sleeps bound in ink, but you can feel its ancient greed stirring with each unpaid installment. The hall's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran financial institutions that were converted for wartime use.}

\exittext{The ledgers balance as Goldenfields' name is spoken. The ink binding loosens, and the village finds freedom in debt forgiveness. The distant echo of the Noon Bell fades with satisfaction, and the hall remembers its former purpose as a place of fair commerce.}
\end{floorbox}

% --- Floor 25 ---
\begin{floorbox}{25 --- The Faith He Broke}
\memoryline{He abandoned a patron's calling mid-rite.}
\tonetags{contrition • chapel • icons • apostasy}
\startposition{Dominant}
\floorclock{[4] --- Omen turns its face; icon drinks breath; faithful arrive demanding answers.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The calling had felt false, the patron's demands unreasonable, the ritual's requirements too restrictive. He told himself he was seeking a more authentic path, that the established religion was corrupt, that his departure was a necessary act of spiritual independence. But the truth was that commitment had become inconvenient, faith had become work, and the easier path was simply to walk away from obligations that no longer served his personal growth. During the Siege of Ecktoria, such abandonments of faith would become common as desperate citizens sought new sources of hope.

\textbf{From the Harmed Party's Perspective:} They had invested their faith in him, seen him as a vessel of divine will, a bridge between the mortal and the sacred. When he abandoned the rite, when he turned his back on the calling that had brought them hope, it felt like a betrayal of their deepest beliefs. Justice would have been dedication---honoring the trust placed in him, completing what he had started, understanding that some commitments transcend personal convenience. Many Utaran faithful faced similar betrayals as their empire's religious institutions collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his abandonment and honor the faith that was placed in him. The lich recognizes that his abandonment of faith mirrored the empire's abandonment of its citizens' spiritual needs.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Chapel half-lit; icons inverted; the air hums with broken devotion. The chapel's design reflects Utaran religious traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Inverted icons bearing Utaran religious markings, abandoned altar with Utaran ceremonial symbols, faith records with Utaran administrative numbering, calling scroll with Utaran religious seals, prayer beads with Utaran spiritual symbols

\textbf{Senses:} The scent of old incense and broken faith; the sound of silent prayers; the feel of divine disappointment; the distant echo of the Noon Bell that once marked time in Utaran religious sites.

\textbf{Adversaries/Agents:} Faithful devotees, icon spirits, calling echoes, apostasy wraiths, divine enforcers, the ghost of a Utaran religious administrator demanding proper conduct.
\end{gmnote}

\handles{Rites (renew vow, DV 2) • Sway (argue theology, DV 3) • Endure (fast, DV 3) • Skirmish (exorcise dissenters, DV 2)}

\textbf{Position Drift:} Renewing raises Position next floor; apostasy drops it.

\textbf{When the Floor Clock Fills:} Omen turns its face in disappointment; icon drinks a PC's breath; faithful arrive demanding answers about the abandoned rite. An inverted icon reveals Utaran religious regulations that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Omen turns its face, withdrawing favor. The withdrawal echoes the loss of divine favor that plagued Utaran religious sites during wartime.}
    \sbitem{2}{Icon drinks a PC's breath, causing Fatigue. The breath carries the essence of Utaran faithful who died when their temples were destroyed during the Siege of Ecktoria.}
    \sbitem{3}{Faithful arrive, demanding explanations. The faithful wear the robes of Utaran religious figures who were forced to abandon their posts during wartime.}
    \sbitem{4+}{The patron's avatar manifests, demanding accountability. The avatar bears the markings of Utaran deities that were suppressed during wartime religious reforms.}
\end{description}

\deckdraw{Draw when omens turn or when the patron's avatar appears---omens of broken faith, debts of abandoned calling, echoes of Utaran religious collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE CONGREGATION OF THE RADIANT DAWN, BETRAYED}
\nameimportance{The congregation had an identity beyond being followers of a faithless leader. Its name connects to Utaran religious traditions, and its betrayal represents the empire's abandonment of its spiritual foundations.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of prayer verses spell the congregation name, with verses bearing Utaran religious text systems
    \item Palimpsest: Congregation name visible only when proper reverence is shown, revealed beneath erased Utaran religious records
    \item Emotional Key: Appears when a PC chooses commitment over convenience, echoing the Utaran values of spiritual dedication
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Prayer verses arranged in congregation initials, with the arrangement matching Utaran religious honorifics
    \item[Actionable Lead:] Rites reveal their full story in church records bearing Utaran religious seals
    \item[Aha! Proof:] Sway work uncovers their individual faith journeys, including references to Utaran religious ceremonies that were cancelled during wartime
\end{description}

\textbf{Extraction Test:} Rites or Sway to restore their faith (on fail, more doubt spreads but reveals congregation details and connections to Utaran religious collapse).

\textbf{Verification \& Use:} Speak the congregation's name while renewing the vow; mispronunciations cause the icons to turn hostile (harm 1) and add a clock segment. Speaking the full name correctly causes the chapel to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Renew (gain Omen: Steady)

\textbf{Restitution:} Apostasy (cut ties; Spade omen follows)

\textbf{Renunciation:} Name a New Path (unique boon keyed to table's Patron themes)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Steady Omen (once: +1d to resist divine disappointment)}
\burden{Spade omen if you apostatize; Ledger - Helped: Radiant Dawn if you renew}
\textbf{Shard Kit:} Prayer bead grants +1d to recognize false faith later; the bead bears Utaran religious markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their faith; the congregation's betrayal represents Utaran's religious collapse
\textbf{Name Ledger Update:} Harmed: Congregation of the Radiant Dawn • Helped: [ ] • Unresolved: [ ]

\entrytext{The chapel sits in half-light, its icons inverted like accusations against a faithless sky. The air hums with the weight of broken devotion, and shadows gather where prayers once illuminated hope. The chapel's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran religious sites that were desecrated during wartime conflicts.}

\exittext{The icons right themselves as the congregation's name is spoken. Light returns to the chapel, and faith finds renewal in honest commitment. The distant echo of the Noon Bell fades with satisfaction, and the chapel remembers its former purpose as a place of proper worship.}
\end{floorbox}

% --- Floor 26 ---
\begin{floorbox}{26 --- The Ward Unkept}
\memoryline{He ignored a binding to chase a rumor.}
\tonetags{contrition • house • mirrors • counting}
\startposition{Dominant}
\floorclock{[5] --- Counting accelerates; mirrors disagree; captive shade grows stronger.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The rumor promised something greater---knowledge that could revolutionize everything, power that could make his current work obsolete. He told himself the binding was temporary, that the shade could wait, that his research was more important than routine maintenance. But the shade had been counting on him, trusting in his commitment, and his absence had allowed entropy to seep in, corruption to take root. In chasing the extraordinary, he had abandoned his ordinary duty. During the Siege of Ecktoria, such abandonments of duty would become common as desperate officials prioritized immediate needs over long-term obligations.

\textbf{From the Harmed Party's Perspective:} They had been promised protection, a safe harbor from the chaos of the world beyond. When their guardian vanished, when the wards grew thin and the boundaries blurred, it felt like being left to drown in a sea of their own fears. Justice would have been reliability---consistent attention to duty, understanding that some responsibilities couldn't be postponed for the sake of ambition. Many Utaran citizens faced similar abandonments as their empire's protective systems collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his neglect and properly tend to the binding he abandoned. The lich recognizes that his neglect of duty mirrored the empire's neglect of its citizens' fundamental needs.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} House with mirrored halls; a captive shade counting in the darkness; ward symbols flicker with fading power. The house's design reflects Utaran residential traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Mirrored halls bearing Utaran architectural markings, captive shade with Utaran magical symbols, ward symbols with Utaran protective markings, binding focus with Utaran ceremonial symbols, counting device with Utaran numerical systems

\textbf{Senses:} The feel of being watched from multiple angles; the sound of obsessive counting; the taste of thinning reality; the distant echo of the Noon Bell that once marked time in Utaran residential areas.

\textbf{Adversaries/Agents:} Mirror wraiths, counting spirits, neglected ward guardians, reality bleeders, shade echoes, the ghost of a Utaran residential administrator demanding proper maintenance.
\end{gmnote}

\handles{Arcana (rebinding, DV 3) • Insight (why he left, DV 2) • Tinker (replace anchor, DV 3) • Skirmish (contain spillover, DV 2)}

\textbf{Position Drift:} Finishing the binding raises Position next floor; freeing the shade drops it.

\textbf{When the Floor Clock Fills:} Counting accelerates to a maddening rhythm; two mirrors disagree about reality, creating paradox; the captive shade grows stronger and more demanding. A flickering ward symbol reveals Utaran residential protection regulations that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Counting accelerates, causing mental stress.}
    \sbitem{2}{Mirror reflections show different truths, creating confusion. The confusion echoes the propaganda that overtook Utaran society during wartime.}
    \sbitem{3}{Two mirrors disagree, causing reality distortions. The distortions echo the breakdown of truth that occurred in Utaran institutions during wartime.}
    \sbitem{4+}{The captive shade fully manifests, demanding release or proper binding. The shade wears the clothes of a Utaran citizen who died when their home's protective wards failed during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when mirrors disagree or when the shade manifests---omens of neglected duty, debts of broken protection, echoes of Utaran residential collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE SHADE OF MARGARET THORNFIELD, NEGLECTED IN DUTY}
\nameimportance{Restoring her identity acknowledges she was more than just a maintenance task. Her name connects to Utaran residential traditions, and her neglect represents the empire's abandonment of its citizens' basic security.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Palimpsest: Her name visible only when proper warding light touches the mirrors, revealed beneath erased Utaran residential records
    \item Syllable Scavenger: Fragments of her name found in the counting sequence bearing Utaran numerical systems
    \item Emotional Key: Appears when a PC chooses responsibility over ambition, echoing the Utaran values of duty and service
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Counting sequence includes numbers that spell her initials, with the sequence matching Utaran residential numbering systems
    \item[Actionable Lead:] Arcana reveals her full story in binding records bearing Utaran magical seals
    \item[Aha! Proof:] Insight uncovers why she needed protection from her own history, including references to Utaran residential security systems that failed during wartime
\end{description}

\textbf{Extraction Test:} Arcana or Insight to understand her needs (on fail, the binding becomes more unstable but reveals her middle name and residential connections).

\textbf{Verification \& Use:} Speak her full name while completing the binding; mispronunciations cause the mirrors to shatter (harm 1) and add a clock segment. Speaking her full name correctly causes the house to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Finish the Binding (boon vs ghosts)

\textbf{Restitution:} Free the Shade (ally later; draw Deck)

\textbf{Renunciation:} Break All Mirrors (scene safer; lose ally)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Proper Binding (once: +1d vs spectral intrusion)}
\burden{Deck draw if you free the shade; Ledger - Helped: Margaret if you finish binding}
\textbf{Shard Kit:} Ward symbol grants +1d to recognize neglected responsibilities later; the symbol bears Utaran protective markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored her name; her neglect represents Utaran's residential collapse
\textbf{Name Ledger Update:} Harmed: Margaret Thornfield • Helped: [ ] • Unresolved: [ ]

\entrytext{The mirrored halls stretch endlessly, each reflection showing a slightly different version of yourself. In the distance, a shade counts obsessively, and ward symbols flicker like dying stars. The house's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran residential buildings that were abandoned during wartime emergencies.}

\exittext{The mirrors align as Margaret's name is spoken. The counting settles into a peaceful rhythm, and the house finds stability in proper care. The distant echo of the Noon Bell fades with satisfaction, and the residence remembers its former purpose as a place of safe harbor.}
\end{floorbox}

% --- Floor 27 ---
\begin{floorbox}{27 --- The Pilgrim's Road}
\memoryline{He extorted pilgrims for passage.}
\tonetags{contrition • gatehouse • tally-rope • oath}
\startposition{Dominant}
\floorclock{[4] --- Saint's bell rings; riot starts; pilgrim echoes grow desperate.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The gatehouse was his responsibility, a position of trust that came with the right to demand reasonable fees for passage. He told himself the tolls were fair, that pilgrims could afford them, that his services were valuable enough to justify the cost. But as the fees grew higher and the demands more unreasonable, as families were turned away for lack of coin, he had become a barrier to faith itself rather than a facilitator of spiritual journey. During the Siege of Ecktoria, such extortions would become common as desperate officials sought resources by any means.

\textbf{From the Harmed Party's Perspectives:} They had saved for years, sacrificed everything, to make this pilgrimage---a sacred journey that was supposed to be about faith, not finance. When they were turned away for lack of coin, when their devotion was measured in copper instead of commitment, it felt like a betrayal of everything the journey was meant to represent. Justice would have been accessibility---ensuring that spiritual seeking wasn't held hostage by material wealth. Many Utaran pilgrims faced similar betrayals as their empire's religious infrastructure collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his extortion and restore the sacred nature of the pilgrimage. The lich recognizes that his exploitation of pilgrims mirrored the empire's exploitation of its citizens' spiritual needs.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Gatehouse with tally-rope and oath-bucket; pilgrim tokens scatter the ground; the air hums with frustrated devotion. The gatehouse's design reflects Utaran religious traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Gatehouse bearing Utaran religious markings, tally-rope with Utaran numerical systems, oath-bucket with Utaran ceremonial symbols, pilgrim tokens with Utaran spiritual markings, sacred way markers with Utaran directional symbols

\textbf{Senses:} The metallic taste of coin and disappointment; the sound of frustrated prayers; the feel of blocked spiritual yearning; the distant echo of the Noon Bell that once marked time in Utaran religious pilgrimage sites.

\textbf{Adversaries/Agents:} Pilgrim echoes, toll enforcers, saint's avatars, frustrated wayfarers, oath guardians, the ghost of a Utaran religious official demanding proper conduct.
\end{gmnote}

\handles{Law (revoke toll, DV 2) • Sway (refund ritual, DV 3) • Trade (pay now, leverage later, DV 3) • Skirmish (stand-down guards, DV 2)}

\textbf{Position Drift:} Refunding raises Position next floor; keeping tolls drops it.

\textbf{When the Floor Clock Fills:} A saint's bell rings in disapproval; a riot starts among frustrated pilgrims; the oath-bucket overflows with broken promises. A pilgrim token reveals Utaran religious regulations on proper pilgrimage conduct that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{A saint's bell rings, marking divine disapproval. The disapproval echoes the loss of divine favor that plagued Utaran religious sites during wartime.}
    \sbitem{2}{Pilgrim tokens multiply, creating confusion. The tokens bear the markings of Utaran pilgrims who died when their journeys were interrupted during the Siege of Ecktoria.}
    \sbitem{3}{A riot starts, with pilgrims demanding passage. The riot echoes the civil unrest that overtook Utaran cities during wartime.}
    \sbitem{4+}{The oath-bucket overflows, unleashing the weight of broken promises. The overflow carries the essence of Utaran pilgrims who died when their sacred journeys were prevented during wartime.}
\end{description}

\deckdraw{Draw when the saint's bell rings or when the oath-bucket overflows---omens of spiritual corruption, debts of blocked faith, echoes of Utaran religious collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE PILGRIMS OF THE SACRED WAY, DENIED PASSAGE}
\nameimportance{They were seekers with a shared spiritual purpose, not just customers to be exploited. Their collective name connects to Utaran religious traditions, and their denial represents the empire's abandonment of its citizens' spiritual freedom.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of sacred way markers spell "pilgrims", with markers bearing Utaran religious directional systems
    \item Palimpsest: Pilgrim name visible only when proper reverence is shown to tokens, revealed beneath erased Utaran religious records
    \item Emotional Key: Appears when a PC chooses faith over profit, echoing the Utaran values of spiritual accessibility
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Sacred way markers arranged in pilgrim formation, with the arrangement matching Utaran religious pilgrimage routes
    \item[Actionable Lead:] Law reveals their full story in pilgrimage records bearing Utaran religious seals
    \item[Aha! Proof:] Sway work uncovers their individual spiritual journeys, including references to Utaran religious pilgrimage traditions that were suspended during wartime
\end{description}

\textbf{Extraction Test:} Law or Sway to understand their quest (on fail, more pilgrims are turned away but reveal way marker details and connections to Utaran religious collapse).

\textbf{Verification \& Use:} Speak the pilgrims' collective name while opening the gate; mispronunciations cause the way markers to turn hostile (harm 1) and add a clock segment. Speaking the full name correctly causes the gatehouse to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Refund (Public Grace)

\textbf{Restitution:} Keep Tolls (resources; Stain)

\textbf{Renunciation:} Turn the Gate (open safe shortcut later)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Pilgrim's Blessing (once: +1d to sacred journeys)}
\burden{Stain if you keep tolls; Ledger - Helped: Pilgrims if you refund}
\textbf{Shard Kit:} Pilgrim token grants +1d to recognize spiritual sincerity later; the token bears Utaran spiritual markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their quest; their denial represents Utaran's religious collapse
\textbf{Name Ledger Update:} Harmed: Pilgrims of the Sacred Way • Helped: [ ] • Unresolved: [ ]

\entrytext{The gatehouse stands before you, its tally-rope marked with the weight of countless tolls. Pilgrim tokens scatter the ground like fallen hopes, and the air hums with frustrated devotion and broken promises. The gatehouse's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran religious pilgrimage sites that were converted for wartime use.}

\exittext{The gate swings open as the pilgrims' name is spoken. The tally-rope unravels, and the way becomes clear for all who seek with honest hearts. The distant echo of the Noon Bell fades with satisfaction, and the gatehouse remembers its former purpose as a place of proper passage.}
\end{floorbox}

% --- Floor 28 ---
\begin{floorbox}{28 --- The Crown of Ashes}
\memoryline{He accepted a crown promised to another.}
\tonetags{contrition • throne room • ash • crowns}
\startposition{Dominant}
\floorclock{[6] --- Ash-snow blinds; spectral court demands oaths; \\ rightful heir's echo rises.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The crown had been within reach, the power so close he could taste it. He told himself the rightful heir was unworthy, that the kingdom needed his vision, his strength, his superior capabilities. But the truth was simpler: ambition had overwhelmed honor, and he had taken what was never his to claim. The other candidate had been chosen by tradition, by the will of the people, by promises made in good faith---and he had cast all of that aside for the sake of personal glory. During the Siege of Ecktoria, such usurpations would become common as desperate officials sought to consolidate power.

\textbf{From the Harmed Party's Perspective:} They had been chosen through proper channels, had earned the right to rule through birth, preparation, and the trust of advisors. When that choice was overturned, when their destiny was stolen by someone more opportunistic, it felt like a theft of their very identity. Justice would have been recognition---honoring the established process, respecting the will of those who had made the choice. Many Utaran officials faced similar usurpations as their empire's political structure collapsed during wartime.

\textbf{Penance Sought:} To acknowledge his usurpation and restore the rightful order he disrupted. The lich recognizes that his usurpation of power mirrored the empire's usurpation of its citizens' right to proper governance.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} Throne room under drifting ash, two crowns on one chair; spectral courtiers watch from shadowy galleries. The throne room's design reflects Utaran governmental traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Throne bearing Utaran governmental markings, ash-drift with Utaran symbolic elements, dual crowns with Utaran royal symbols, succession records with Utaran administrative numbering, royal regalia with Utaran ceremonial markings

\textbf{Senses:} The taste of ash and ambition; the sound of spectral whispers; the feel of contested authority; the distant echo of the Noon Bell that once marked time in Utaran governmental buildings.

\textbf{Adversaries/Agents:} Ash wraiths, spectral courtiers, rightful heir's echo, succession guardians, crown spirits, the ghost of a Utaran governmental official demanding proper succession.
\end{gmnote}

\handles{Sway (abdicate, DV 2) • Duel (trial-by-steel, DV 3) • Evidence (prove the promise, DV 3) • Arcana (split the crown, DV 2)}

\textbf{Position Drift:} Abdicating raises Position next floor; ruling by fear drops it.

\textbf{When the Floor Clock Fills:} Ash-snow blinds all present; spectral court demands new oaths of loyalty; the rightful heir's echo rises to claim what was stolen. A succession record reveals Utaran governmental regulations on proper succession that were suspended during wartime emergencies.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Ash-snow blinds, causing disorientation. The blindness echoes the confusion that overtook Utaran governmental institutions during wartime.}
    \sbitem{2}{Spectral courtiers whisper accusations. The accusations echo the political purges that plagued Utaran officials during wartime.}
    \sbitem{3}{Succession records catch fire, threatening to destroy the truth. The destruction echoes the burning of Utaran governmental archives during wartime conflicts.}
    \sbitem{4+}{The rightful heir's echo fully manifests, demanding justice. The echo wears the robes of a Utaran official who died when their legitimate authority was usurped during the Siege of Ecktoria.}
\end{description}

\deckdraw{Draw when the heir's echo rises or when succession records are threatened---omens of illegitimate power, debts of stolen destiny, echoes of Utaran governmental collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{PRINCE/COUNCILOR ALARIC STERLING, RIGHTFUL HEIR USURPED}
\nameimportance{Restoring his full title acknowledges his legitimate claim beyond being an obstacle to be removed. His title connects to Utaran governmental traditions, and his usurpation represents the empire's abandonment of its citizens' right to proper leadership.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Acrostic: Initial letters of succession law texts spell his name, with texts bearing Utaran governmental legal systems
    \item Palimpsest: His name visible only when proper regalia is placed on the throne, revealed beneath erased Utaran governmental records
    \item Emotional Key: Appears when a PC chooses legitimacy over personal gain, echoing the Utaran values of proper governance
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] Succession texts arranged in his initials, with the arrangement matching Utaran governmental honorifics
    \item[Actionable Lead:] Evidence reveals his full claim in royal archives bearing Utaran governmental seals
    \item[Aha! Proof:] Sway work uncovers the court's true opinion of him, including references to Utaran governmental succession protocols that were abandoned during wartime
\end{description}

\textbf{Extraction Test:} Evidence or Sway to validate his claim (on fail, more evidence is destroyed but reveals his middle name and governmental credentials).

\textbf{Verification \& Use:} Speak his full name while placing proper regalia; mispronunciations cause the crowns to fight each other (harm 1) and add a clock segment. Speaking his full title correctly causes the throne room to resonate with the tone of the Noon Bell.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Abdicate (DV -1 later vs authority)

\textbf{Restitution:} Rule by Fear (Desperate tag sticks 1 floor)

\textbf{Renunciation:} Share Crown (unlock ally later)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Legitimacy's Shield (once: +1d vs authority challenges)}
\burden{Desperate tag if you rule by fear; Ledger - Helped: Alaric if you abdicate}
\textbf{Shard Kit:} Royal regalia fragment grants +1d to recognize legitimate authority later; the fragment bears Utaran governmental markings
\textbf{Links:} Reduces Final Persuasion DV by 1 if you honored his name; his usurpation represents Utaran's governmental collapse
\textbf{Name Ledger Update:} Harmed: Alaric Sterling • Helped: [ ] • Unresolved: [ ]

\entrytext{The throne room lies under a constant drift of ash, two crowns resting on the seat like accusations against each other. In the galleries, spectral courtiers watch with eyes that remember betrayal, and the air hums with contested legitimacy. The throne room's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran governmental buildings that were destroyed during wartime conflicts.}

\exittext{The ash settles as Alaric's name is spoken. One crown fades away while the other settles properly on the throne, and the court finds peace in restored order. The distant echo of the Noon Bell fades with satisfaction, and the throne room remembers its former purpose as a place of proper governance.}
\end{floorbox}

% --- Floor 29 ---

\begin{floorbox}{29 --- The Night of Names}
    \memoryline{He erased a persecuted lineage to hide himself.}
    \tonetags{contrition • scriptorium • fading names • persecution}
    \startposition{Dominant}
    \floorclock{[5] --- Names blur from ledgers; family ghosts rise; inquisitor's seal activates.}
    
    \begin{gmnote}{Extended Lore (For GM Eyes)}
    \textbf{From the Lich's Perspective:} The family name had become a liability---a target for those seeking to punish past sins by proxy. He told himself he was protecting them, that erasure was mercy, that disappearing the lineage would shield them from persecution. But the truth was that shame had driven him to deny his own blood, to pretend the connection had never existed rather than face the consequences of association. During the Siege of Ecktoria, such erasures of identity would become common as desperate families sought to avoid being targeted for their connections.
    
    \textbf{From the Harmed Party's Perspective:} They had done nothing wrong---their only crime was bearing a name that someone else had made infamous. When their identity was stripped away, when records were falsified and history was rewritten, it felt like a death of the soul. Justice would have been protection---keeping family connections intact, finding safe ways to shelter without destroying heritage. Many Utaran families faced similar erasures as their empire's persecution systems targeted entire lineages during wartime.
    
    \textbf{Penance Sought:} To acknowledge his erasure of their identity and restore what he took from his own flesh and blood. The lich recognizes that his denial of family mirrored the empire's denial of its own diverse heritage.
    \end{gmnote}
    
    \begin{gmnote}{Environment \& Manifestations}
    \textbf{Space:} Scriptorium of fading names; half-written ledgers; family portraits with faces blurred by time and neglect. The scriptorium's design reflects Utaran administrative traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.
    
    \textbf{Anchors:} Fading ledgers bearing Utaran administrative markings, blurred portraits with Utaran noble crests, erasure tools with Utaran bureaucratic symbols, family seals with Utaran heraldic markings, inquisitor's stamp with Utaran legal seals
    
    \textbf{Senses:} The scent of old parchment and forgotten memories; the sound of names being whispered then lost; the feel of identity slipping away; the distant echo of the Noon Bell that once marked time in Utaran family halls.
    
    \textbf{Adversaries/Agents:} Name-thieves, family echoes, inquisitor spirits, erasure wraiths, identity hunters, the ghost of a Utaran administrative official demanding proper record-keeping.
    \end{gmnote}
    
    \handles{Evidence (restore records, DV 3) • Arcana (reweave names, DV 2) • Skirmish (fight name-thieves, DV 3) • Sway (appease inquisitor, DV 2)}
    
    \textbf{Position Drift:} Restoring names raises Position next floor; burning records drops it.
    
    \textbf{When the Floor Clock Fills:} Names blur completely from all ledgers; family ghosts rise demanding recognition; inquisitor's seal activates, summoning judgment. A blurred portrait reveals Utaran identity protection regulations that were suspended during wartime emergencies.
    
    \textbf{SB Spends (Revelation-Forward)}
    \begin{description}[leftmargin=*, style=unboxed]
        \sbitem{1}{Names blur from ledgers, causing confusion. The blurring echoes the systematic removal of Utaran family names during wartime purges.}
        \sbitem{2}{Family ledgers show more gaps, indicating wider erasure. The gaps echo the systematic removal of Utaran noble families during wartime purges.}
        \sbitem{3}{A PC risks losing their past to the same name-stealing magic. The magic carries the essence of Utaran citizens who had their identities erased during the Siege of Ecktoria.}
        \sbitem{4+}{Inquisitors manifest, demanding to know hidden identities. The inquisitors wear the uniforms of Utaran security forces who conducted identity purges during wartime.}
    \end{description}
    
    \deckdraw{Draw when names blur or when inquisitors arrive---omens of identity theft, debts of denied heritage, echoes of Utaran social collapse.}
    
    \begin{gmnote}{The Name-as-Key}
    \keyname{THE STERLING LINE, ERASED IN FEAR}
    \nameimportance{The family had an identity beyond being a source of shame. Their name connects to Utaran noble traditions, and their erasure represents the empire's systematic destruction of its own heritage.}
    
    \textbf{How the Name Is Hidden (choose 1--2):}
    \begin{itemize}[leftmargin=*]
        \item Palimpsest: Family name visible only when tears fall on erasure marks, revealed beneath erased Utaran noble records
        \item Syllable Scavenger: Fragments of name found in faded ledger entries bearing Utaran administrative numbering systems
        \item Emotional Key: Appears when a PC chooses acceptance over shame, echoing the Utaran values of honoring one's heritage
    \end{itemize}
    
    \textbf{Clue Ladder:}
    \begin{description}[leftmargin=*, style=unboxed]
        \item[Gentle Nudge:] Faded ledger entries spell family initials, with the entries matching Utaran noble record systems
        \item[Actionable Lead:] Evidence reveals full family history in hidden records bearing Utaran noble seals
        \item[Aha! Proof:] Arcana work uncovers the true extent of the erasure, including references to Utaran noble families that were systematically removed during wartime purges
    \end{description}
    
    \textbf{Extraction Test:} Evidence or Arcana to restore family honor (on fail, more names are lost but reveal family details and connections to Utaran noble houses).
    
    \textbf{Verification \& Use:} Speak the family name while restoring records; mispronunciations cause more names to fade (harm 1) and add a clock segment. Speaking the full name correctly causes the scriptorium to resonate with the tone of the Noon Bell.
    \end{gmnote}
    
    \begin{resolutionbox}
    \textbf{Mercy:} Restore Line (honor family memory)
    
    \textbf{Restitution:} Burn It All (safety now, haunting later)
    
    \textbf{Renunciation:} Sign Your Own True Name (gain Anchor boon)
    \end{resolutionbox}
    
    \textbf{Outcomes \& Carry-Forward}
    \boon{Family Anchor (once: immunity to identity theft)}
    \burden{Future haunting if you burn it all; Ledger - Helped: Sterling Line if you restore}
    \textbf{Shard Kit:} Family ledger page grants +1d to recognize erased histories later; the page bears Utaran noble markings
    \textbf{Links:} Reduces Final Persuasion DV by 1 if you honored their memory; the family's erasure represents Utaran's social collapse
    \textbf{Name Ledger Update:} Harmed: Sterling Family Line • Helped: [ ] • Unresolved: [ ]
    
    \entrytext{The scriptorium's pages shimmer with half-finished names that fade as you look at them. Ancient ledgers show gaps where families once flourished, and the air hums with the whispers of forgotten identities seeking remembrance. The scriptorium's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran noble archives that were systematically destroyed during wartime purges.}
    
    \exittext{The pages stabilize as the family name is spoken. Letters flow into place like returning souls, and the scriptorium finds peace in restored memory. The distant echo of the Noon Bell fades with satisfaction, and the archive remembers its former purpose as a place of proper record-keeping.}
    \end{floorbox}
    

% --- Floor 30 ---
\begin{floorbox}{30 --- The Last Light}
\memoryline{He chose lichdom to pay for everything.}
\tonetags{contrition • study • candles • choice}
\startposition{Dominant if you showed mercy/restitution on ≥10 floors; Dominant otherwise. Any violent approach starts Desperate.}

\begin{gmnote}{Extended Lore (For GM Eyes)}
\textbf{From the Lich's Perspective:} The weight of his sins had become unbearable---a mountain of guilt that crushed every attempt at peace. He told himself that lichdom was penance, that eternal existence would give him infinite time to atone, that by refusing death he was refusing to escape from the consequences of his actions. But the truth was that he was afraid---afraid of judgment, afraid of oblivion, afraid of facing the final accounting of his life. In choosing undeath, he had chosen to postpone rather than confront. During the Siege of Ecktoria, such refusals to face consequences would become common as desperate officials sought to avoid accountability for their actions.

\textbf{From the Harmed Party's Perspective:} They had wanted many things---justice, acknowledgment, closure, the simple satisfaction of knowing their pain mattered to someone who had caused it. When he chose lichdom, when he opted for eternal self-torment over the possibility of peace, it felt like one final betrayal. Justice would have been courage---facing the consequences, accepting the end that comes to all mortals, understanding that some debts can only be paid with the ultimate sacrifice. Many Utaran citizens faced similar refusals to accept consequences as their empire's leaders sought to avoid accountability for the collapse.

\textbf{Penance Sought:} To acknowledge that his penance was complete and choose what comes next. The lich recognizes that his refusal to accept mortality mirrored the empire's refusal to accept its own limitations and mortality.
\end{gmnote}

\begin{gmnote}{Environment \& Manifestations}
\textbf{Space:} A quiet study of candles. Thirty flames, each the height of a floor you climbed. The lich sits unarmed; the air is bright with names unsaid. The study's design reflects Utaran scholarly traditions---triple-course river-stone construction with laurel-circle decorative elements and Utaran civic symbols.

\textbf{Anchors:} Candle study bearing Utaran academic markings, thirty flames arranged like Utaran memorial candles, unarmed lich with Utaran noble features, name ledger with Utaran administrative numbering, final choice focus with Utaran philosophical symbols

\textbf{Senses:} The scent of burning wax and old regrets; the warmth of thirty separate memories; the weight of final choice; the clear, pure tone of the Noon Bell that finally rings true after centuries of discord.

\textbf{Adversaries/Agents:} Memory echoes, guilt manifestations, thirty separate betrayals made manifest, the weight of accumulated sin, the ghost of Utaran philosophers demanding proper acceptance of mortality.
\end{gmnote}

\handles{Sway/Presence (argue, DV varies by prior choices) • Insight (name what he fears to lose, DV varies) • Evidence (lay the ledger, DV varies) • Rites (open a door beyond, DV varies)}

\textbf{Position Drift:} Nonviolent approaches improve Position; violence worsens it dramatically.

\textbf{When the Floor Clock Fills:} All candles flare simultaneously, forcing a final choice; memory echoes rise to testify; the lich's true form flickers between life and undeath. The flaring echoes the final moments of Utaran civilization when all its institutions collapsed simultaneously.

\textbf{SB Spends (Revelation-Forward)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Candles gutter images of those you saved. The images show Utaran citizens who were helped during the Siege of Ecktoria by choices made in earlier floors.}
    \sbitem{2}{Memory echoes rise to testify about specific betrayals. The echoes wear the clothes of Utaran citizens who died during the empire's collapse.}
    \sbitem{3}{A beloved echo begs you to forbid forgiveness. The echo wears the robes of a Utaran philosopher who died trying to preserve the empire's moral foundations.}
    \sbitem{4+}{The lich's form becomes unstable, threatening to collapse entirely. The instability echoes the final collapse of Utaran civilization when its institutions could no longer sustain themselves.}
\end{description}

\deckdraw{Draw when memory echoes testify or when the lich's form becomes unstable---omens of final judgment, debts of ultimate choice, echoes of Utaran civilizational collapse.}

\begin{gmnote}{The Name-as-Key}
\keyname{THE THIRTIETH NAME, THE CHOICE OF UNDEATH}
\nameimportance{The final name is the choice itself---the acknowledgment that he defined himself by his refusal to end. This choice connects to the fundamental question of how Utaran civilization chose to face its own mortality and limitations.}

\textbf{How the Name Is Hidden (choose 1--2):}
\begin{itemize}[leftmargin=*]
    \item Emotional Key: Appears when the party has shown sufficient mercy, reflecting the Utaran philosophical tradition of mercy as the highest virtue
    \item Ledger Lock: Must be written with all thirty previous names acknowledged, representing the Utaran legal tradition of proper record-keeping
    \item Syllable Scavenger: The choice is spelled out in the first letter of each floor's key name, following the Utaran scholarly tradition of hidden meanings in texts
\end{itemize}

\textbf{Clue Ladder:}
\begin{description}[leftmargin=*, style=unboxed]
    \item[Gentle Nudge:] The thirtieth candle burns differently from the others, with a flame that echoes the pure tone of the Noon Bell that marked proper time in Utaran civilization
    \item[Actionable Lead:] Insight reveals that the final choice is his to make, but only after acknowledging the full weight of Utaran civilization's collapse that his actions contributed to
    \item[Aha! Proof:] All previous names must be spoken to unlock the final truth, representing the Utaran philosophical belief that understanding requires acknowledging all aspects of a situation
\end{description}

\textbf{Extraction Test:} The final test is not mechanical but moral---can the party convince him that his penance is complete and that he can face the same mortality that claimed Utaran civilization?

\textbf{Verification \& Use:} Speak the thirtieth name while making the final choice; mispronunciations cause all candles to extinguish (harm 2) and force a violent resolution. Speaking the name correctly causes the study to resonate with the clear, pure tone of the Noon Bell that finally rings true after centuries of discord.
\end{gmnote}

\begin{resolutionbox}
\textbf{Mercy:} Continue Penance (The lich seals himself deeper. Grant Fetter's Boon)

\textbf{Restitution:} Find Peace (He passes on; a Soul-Light remains)

\textbf{Renunciation:} Join the Living (He returns in frail flesh; a Mortal Oath binds him)
\end{resolutionbox}

\textbf{Outcomes \& Carry-Forward}
\boon{Soul-Light (1/arc: remove a Desperate → Dominant shift for one scene)}
\burden{Fetter's Boon requires the GM to bank +1 SB; Ledger - Helped: All previous if you find peace}
\textbf{Shard Kit:} Final candle grants permanent +1d to recognize true contrition; the candle bears the markings of Utaran memorial traditions
\textbf{Links:} The final choice determines the campaign's legacy and any future encounters; the choice reflects how one faces the same mortality that claimed Utaran civilization
\textbf{Name Ledger Update:} Harmed: All previous • Helped: Those you convinced • Unresolved: Those you couldn't reach

\entrytext{Thirty candles burn in perfect stillness, each one representing a floor climbed, a memory faced, a betrayal acknowledged. The lich sits before them, unarmed and waiting, his eyes reflecting the light of every name you have carried this far. The study's construction features triple-course river-stone walls and laurel-circle motifs, reminiscent of Utaran philosophical institutions that sought to understand the nature of existence and mortality.}

\exittext{The final candle's flame dances as the thirtieth name is spoken. Whether it settles into peaceful darkness, continues its eternal vigil, or kindles anew with mortal hope depends on the choice made in this quiet study where all debts come due. The Noon Bell finally rings clear and true, its pure tone echoing across the dungeon like the sound of proper justice being served---a sound that Utaran civilization never heard before its collapse, but which its memory now acknowledges was always there, waiting to be rediscovered.}
\end{floorbox}

\subsection*{Combat Clocks (Violence Initiated)}
\emph{If players initiate violence, introduce the following clocks immediately and advance them based on aggression and SB spends.}

\paragraph{Lich's Phylactery Pulse Clock [4]}
Represents the Lich's core life force and the dungeon's reaction to violence. \emph{Starts at 0. Advances via SB spends or direct harm dealt to the Lich.}
\begin{itemize}
    \item \textbf{When it fills 1:} The Lich becomes Immune to Harm 1. (\emph{Narrative: His form shimmers, absorbing minor blows}).
    \item \textbf{When it fills 2:} The Lich may ignore one Desperate result per turn (treat as Dominant) OR gain +1 Effect on one action. (\emph{Narrative: Power drawn from his eternal anchor}).
    \item \textbf{When it fills 3:} The Lich may immediately take an extra action once. (\emph{Narrative: Surge of necrotic energy}).
    \item \textbf{When it fills 4:} Reset to 0. Trigger Major Backlash (Fate/Death element - options include: Immediate, unavoidable Harm 2 to all aggressors; Severe environmental collapse affecting the entire study; Summon a powerful, bound memory-echo as an ally). (\emph{Narrative: The dungeon itself lashes out}).
\end{itemize}

\paragraph{Lich's True Form Manifestation Clock [6]}
Represents the Lich shedding his penitent facade.
\begin{itemize}
    \item \emph{Starts at 0. Advances via SB spends, direct harm, or failed non-violent resolution attempts once violence starts.}
    \item \textbf{When it fills 2:} The Lich's stats increase (DV +1 for actions against him, +1 Effect for his actions). His appearance subtly shifts (eyes glow, shadows deepen).
    \item \textbf{When it fills 4:} The Lich's stats increase further (DV +1, +1 Effect). He gains access to potent, AoE magical attacks (e.g., \emph{Circle of Woe} - targets all PCs, DV 4, deals Harm 1 and inflicts Condition: \emph{Shaken} on failure/partial).
    \item \textbf{When it fills 6:} The Lich is in his full, terrible glory. He may now act with Dominant Position against all opponents. He can invoke a Rite or powerful magic effect once without cost (e.g., \emph{Unmaking Word} - target one PC, DV 5, Severe consequence like Incapacitation or Major Item Destruction). Treat further resistance as Desperate for the party.
\end{itemize}

\paragraph{Party Dissolution Clock [4]}
Represents the party being overwhelmed and defeated.
\begin{itemize}
    \item \emph{Starts at 0. Advances when a PC takes Harm, is reduced to 0 Stress/Strain, fails a crucial roll under pressure from the Lich, or when a major environmental effect impacts them.}
    \item \textbf{When it fills 1:} One PC must mark Fatigue.
    \item \textbf{When it fills 2:} One PC takes automatic Harm 1.
    \item \textbf{When it fills 3:} One PC takes automatic Harm 2.
    \item \textbf{When it fills 4:} The party is defeated/overwhelmed. Assume TPK or capture/enslavement (triggering a different kind of ``penance'' - perhaps being bound into service or trapped in a memory loop).
\end{itemize}

\subsection*{GM Guidance for Combat}
If players initiate violence:
\begin{enumerate}
    \item Immediately shift the scene's starting Position to \textbf{Desperate}.
    \item Introduce the \textbf{Lich's Phylactery Pulse Clock [4]} and \textbf{Lich's True Form Manifestation Clock [6]}. Place them visibly.
    \item As violence escalates (attacks are made, SB is spent, the Lich is harmed), advance these clocks according to the rules above and the SB spend menu.
    \item Introduce the \textbf{Party Dissolution Clock [4]}. Advance this whenever the party takes significant hits, fails rolls under pressure, or is affected by the Lich's escalating powers.
    \item Describe the Lich's transformation and the environment's reaction vividly. He should feel \emph{wrong}, \emph{immense}, and \emph{inevitable} once unleashed.
    \item Use SB spends liberally to introduce complications, environmental hazards, and to advance the new clocks, reflecting the sheer weight of unleashing this kind of power.
    \item The goal is for the players to realize, very quickly, that this path leads to a swift and total defeat, reinforcing that the \emph{true} challenge lies in the moral and narrative resolution, not martial combat.
\end{enumerate}

% Penitent Lich Patron
\patrontitle{The Penitent Lich --- Runekeeper Patron}

A Patron of names kept, vows witnessed, and debts made visible. Power flows through confession, restitution, and the binding force of true names.

\textbf{What the Penitent Lich Is}

Once mortal, now an anchor of memory. He refuses oblivion until every name he harmed is spoken correctly and set in its rightful place. He grants power to those who lift names, keep watches, and choose mercy first.

\textbf{Portfolio:} true names • oaths • remembrance • wards • restitution

\textbf{Symbols:} a guttering candle; a ledger ribbon; a ring that won't warm; a mirror shard.

\textbf{Favored Deeds:} return a name to the world; bury the unburied; witness a vow without centering yourself; refuse an easy cruelty.

\subsection{Tenets (Keep These)}
\tenet{1}{Name the Harmed: When you act, speak who is owed---not what you want.}
\tenet{2}{Mercy Before Might: Try a nonviolent avenue before you draw steel or spell.}
\tenet{3}{Watches Are Kept, Not Owned: If you take a ward, keep it---or find a worthy keeper.}
\tenet{4}{Debt Is Not Silence: Confess, repair, and let the record show it.}
\tenet{5}{No Erasures: Do not profit by deleting names, histories, or lineages.}

\subsection{Runekeeper: What You Get}

When you swear to the Penitent Lich as Runekeeper, you learn to inscribe small, portable bindings called runes. Each rune is a single-use inscription prepared during downtime or a quiet scene.
\begin{itemize}
    \item \textbf{Preparing Runes:} In downtime, prepare 2 runes from your list; carry at most 3 at a time. You may overwrite an unspent rune by marking Obligation to the Patron.
    \item \textbf{Invoking Runes:} Invoking a rune is usually Dominant and takes a quick action unless the scene is in chaos (then Dominant).
    \item \textbf{Interacting with Position/DV:} Runes usually raise Position one step or give +1 effect towards their stated sphere.
\end{itemize}

\subsection{Devotions \& Boons}

Track your standing with the Penitent Lich as Favor [0--5]. Increase Favor by naming the harmed, choosing restitution, keeping hard watches, or completing a rite of remembrance. Decrease Favor by erasing names, refusing confession, or striking down a pleading foe without terms.

\devotion{0}{Initiate of the Ledger}
\rite{Patron Gift --- Rune: Mark of Witness}{Chalk or ash mark that steadies Position once per scene when you defend the powerless or truth.}
\rite{Prayer---Soft Candle}{Once per session, dim a scene's hostility: -1 segment from an active Doom/Alert/Ward sub-clock after you speak who is being protected.}

\devotion{1}{Keeper of the Door}
\rite{Rune: Threshold}{Place on a doorway/line. For one scene, allies crossing it get +1d to resist separation or ambush.}
\rite{Boon---Confessor's Ears}{In a social exchange where you admit a fault, gain +1d to read motive (Insight-type handles) this scene.}

\devotion{2}{Binder of Names}
\rite{Rune: True Name}{Inscribe a true name on cloth/metal. For one action that targets the named (aid or oppose), choose: +1 effect or steady Position. If used to harm, mark Stain unless you offered terms first.}
\rite{Boon---Ledger Glimpse}{Once per session, ask the GM: ``What would restitution look like here?'' Gain a concrete, actionable step.}

\devotion{3}{Warden of Restitution}
\rite{Rune: Redress}{Break this seal to clear 2 segments from a Debt/Obligation clock you owe---but you must immediately perform a small, public act of repair.}
\rite{Boon---Vigil}{When you keep watch over a warded thing for a full scene, everyone on your side begins the next scene one Position higher.}

\devotion{4}{Litany-Bearer}
\rite{Rune: Litany}{Speak a short list of rightful names. For the next minute, your side treats the scene's first Desperate roll as Dominant. Consumes one distinct name from your Name Ledger.}
\rite{Boon---Mercy Carried Forward}{When you choose a nonviolent exit in a grievous scene, remove 1 segment from a Memory/Alert clock.}

\devotion{5}{Keeper of the Last Light}
\rite{Rite: The Thirty-Spoken}{If your Ledger holds 30 true names from a single penitential place, you may resolve a confrontation with persuasion instead of violence; on success, also refresh one spent rune.}
\rite{Boon---Absolution's Edge}{Once per arc, declare immunity to fear/charm for your party in a scene where you speak for the harmed.}

\subsection{Runes (Runekeeper List)}

Pick from these when you prepare:
\begin{itemize}
    \item Witness: Mark a surface; your side gets +1 effect to protect witnesses or evidence this scene.
    \item Threshold: See Favor 1.
    \item True Name: See Favor 2.
    \item Redress: See Favor 3.
    \item Peace: Anyone who voluntarily sheaths a weapon within arm's reach of this rune gains +1d to their next social action in this scene.
    \item Rest: Set by a body or grave; undead entering must test (DV 2). On fail, they hesitate and reveal who they were.
    \item Atonement: Place on an object tied to harm; returning it to its rightful keeper steadies Position for your group and creates a minor Boons token.
\end{itemize}

\subsection{Rites \& Invocations}
\rite{Candle of Remembrance (scene)}{Light a small candle and speak a harmed name. For this scene, you may re-roll 1s once on a social or investigation action that centers that person.}
\rite{Seal of Safe Passage (scene)}{Trace a circle and recite three names you have helped. For the next crossing (gate, bridge, alley), reduce the first Ward/Alert sub-clock you trigger by 1.}
\rite{Confessor's Vigil (downtime)}{Keep watch at a threshold named in your Ledger. Clear 1 stress/strain and prepare 1 extra rune next downtime.}

\subsection{Compels \& Taboos}
\compel{Offer a boon or clear 1 segment of an impending trouble if the Runekeeper attempts a nonviolent option first. Refusing marks Stain or loses 1 Favor (player's choice).}
\taboo{Erase a name for gain; profit from falsified records; abandon a sworn watch without successor; kill a pleading foe without terms.}

\subsection{Offerings \& Obligations}
\offering{Restore a headstone or ledger page.}
\offering{Return a stolen token (ring, book, oath-stone).}
\offering{Keep a night's watch for someone who cannot.}
\offering{Publicly confess a relevant fault and make a concrete repair.}

\obligation{Mark Obligation [4] when you take aid without an offering. Fulfilling a listed offering clears 2 segments.}

\synergy{If your party earned the Glaive of the Thirty, you may speak a Ledger name while guarding or striking to gain +1 effect or steady Position (once/scene). Re-using the same name before the next arc marks Stain.}

\subsection{Using This Patron in the Lich's Dungeon}
\begin{itemize}
    \item \textbf{Name Ledger:} Keep three columns: Harmed / Helped / Unresolved. Moving names from Unresolved → Helped is how you organically earn Favor.
    \item \textbf{SB Spending Guidance:} Spend SB to reveal truths, tighten oaths, or summon witnesses, not just to spike harm.
    \item \textbf{Nonviolent First:} With this Patron, players should feel tangible benefits for choosing mercy---Position bump, DV hints, and clock relief.
\end{itemize}

\subsection{Example Invocations}
\begin{itemize}
    \item ``By the breath that warmed ALYSSA WHITE-LILY's ring, let no vow be mocked while this mark holds.''
    \item ``Threshold, keep---the harmed may pass before the hungry.''
    \item ``Ledger, open---show me what restitution looks like, not what is easy.''
\end{itemize}

\subsection{GM Notes (Quick)}
\begin{itemize}
    \item Favor should move. Reward visible repair. Make taboos costly but recoverable through clean acts.
    \item When in doubt, let a rune raise Position or grant +1 effect toward mercy, truth, or watchkeeping rather than raw damage.
\end{itemize}

% Penitential Corruption
\section{Penitential Corruption --- Tables \& Rules}

Use this when the dungeon's necromancy stains the soul, when names are erased, or when the party spurns mercy. It integrates with Position, DV, SB, the Deck of Consequences, and the Name Ledger.

\corruptiontitle{How Corruption Works}
\begin{itemize}
    \item \textbf{When to Roll:} Gain Minor Corruption on small cruelties (breaking etiquette of remembrance, exploiting grief, petty desecrations). Gain Major Corruption for taboos (erasing a name, killing a pleading foe without terms, shattering a ward, stealing a ritual for profit, phylactery desecration without rites).
    \item \textbf{Track:} Each PC keeps a Corruption [0--6] track. Each roll adds +1 Corruption first, then you roll on the relevant table. Some results add more.
    \item \textbf{Stain vs. Corruption:} Stain is a narrative tag that can clear through play; Corruption is a mechanical track with thresholds below.
    \item \textbf{GM SB Hook:} In scenes of cruelty or erasure, the GM may spend 2 SB to force a Minor Corruption roll (once/scene).
    \item \textbf{Name Ledger Hook:} At Corruption ≥3, you cannot move any name from Unresolved → Helped without public restitution.
\end{itemize}

\threshold{3}{Choose a Mark (visible stigma). Deception about your past suffers DV +1. You cannot exceed Favor 3 with the Penitent Lich until you reduce to ≤2.}
\threshold{6}{You are Oath-Bound Revenant until cleansing (see below): all violent actions start Desperate; Glaive of the Thirty (if held) goes dormant; a personal Haunt follows you as a GM tool.}

\textbf{Cleansing Corruption (pick/stack as fiction allows)}
\begin{itemize}
    \item Public Confession + Concrete Restitution: -1
    \item Bury the Unburied / Restore a Headstone or Ledger Page: -1
    \item Keep a Named Vigil (one full scene) at a relevant threshold: -1
    \item Return a Stolen Token to its rightful keeper: -1
    \item Restore a Lineage / Undo an Erasure: -2
    \item The Thirty-Spoken (leader only): set Corruption to 0
\end{itemize}

\corruptiontitle{Table A --- Minor Corruption (d12)}
Roll when you take a petty cruelty or minor desecration.
\begin{enumerate}
    \item Cold Breath: Your breath steams in warm rooms. The first time you draw a weapon each scene, tick Alert +1 segment.
    \item Ink-Fingers: Black ink stains your hands. Evidence-type actions suffer DV +1 unless you start by admitting a relevant fault.
    \item Wrong-Whisper: Echoes misname you. Sway vs any harmed party is -1d unless you begin with acknowledgment; on acknowledgment, steady Position.
    \item Guttering Candle: Nearby flames flutter toward you. When you lie about the past, immediately draw the Deck of Consequences.
    \item Numb Grip: Fingers won't warm. Your first delicate action each scene (Tinker/Theft/etc.) starts Dominant unless you carry a token of the harmed.
    \item Cold Ring: A band bites your finger. If you break a promise this scene, tick a Memory Collapse [4] sub-clock +1.
    \item Skittering Shadow: Your shadow jitters. Stealth starts Dominant unless you spoke a true name today.
    \item Frosted Mirror: Your reflection fogs first. When acting from self-deception, DV +1; if you speak your true name, ignore this for the scene.
    \item Ash in Pocket: Coin spent for selfish gain becomes an IOU; tick Obligation +1 segment (to anyone harmed by the scene).
    \item Blurred Margins: Names blur at the edges of pages. Once/scene the GM may ask, ``Whose name is missing?'' If you cannot answer, Position -1.
    \item Hollow Step: Floors creak beneath you. Endure-type resists cost +1 stress/strain unless defending the powerless.
    \item Bleeding Ledger: When you destroy a record, mark Stain and roll Minor Corruption again.
\end{enumerate}

\corruptiontitle{Table B --- Major Corruption (d12)}
Roll when you violate a taboo or commit a grievous erasure.
\begin{enumerate}
    \item Counting Shadow: Your shadow counts softly. At Dusk, accept a Compel to attempt a nonviolent option first or mark Stain.
    \item The Unlit Candle: A thirty-first candle won't catch. Undead prefer you as a target, but you gain +1 effect on Rites after confession.
    \item Tongue of Ash: Mispronouncing a true name adds +1 segment to Persuasion/Memory Collapse and locks that name until restitution.
    \item Name-Scar: Healed wounds etch a name you can't say. Choose an NPC's name you are unable to speak until you perform a fitting repair.
    \item Frosted Vows: Your touch chills vows. Allies start the next social scene Dominant unless you consecrate with Rune: Witness.
    \item Mirror-Shard Soul: On being reduced to 0, you split into two conflicting drives for one scene (GM tool) until healed by a Vigil.
    \item Debtor's Sigil: Mark Obligation [4] to the Patron. Until cleared, you cannot invoke the same Litany or Glaive name twice per arc.
    \item Name-Eater: On a critical failure, one random name moves from Helped → Unresolved; gain +1d to your immediate follow-up and mark Stain.
    \item Ash Crown: Authority senses your usurpations: DV +1 vs authority until you abdicate something tangible.
    \item Counting Bells: Time slips. Each long action you take also ticks Alert/Ward +1 unless a harmed witness is present.
    \item Empty Pall: Your healing helps half unless accompanied by a spoken apology tied
\end{floorbox}

% Penitent Lich Patron
\patrontitle{The Penitent Lich --- Runekeeper Patron}

A Patron of names kept, vows witnessed, and debts made visible. Power flows through confession, restitution, and the binding force of true names.

\textbf{What the Penitent Lich Is}

Once mortal, now an anchor of memory. He refuses oblivion until every name he harmed is spoken correctly and set in its rightful place. He grants power to those who lift names, keep watches, and choose mercy first.

\textbf{Portfolio:} true names • oaths • remembrance • wards • restitution

\textbf{Symbols:} a guttering candle; a ledger ribbon; a ring that won't warm; a mirror shard.

\textbf{Favored Deeds:} return a name to the world; bury the unburied; witness a vow without centering yourself; refuse an easy cruelty.

\subsection{Tenets (Keep These)}
\tenet{1}{Name the Harmed: When you act, speak who is owed---not what you want.}
\tenet{2}{Mercy Before Might: Try a nonviolent avenue before you draw steel or spell.}
\tenet{3}{Watches Are Kept, Not Owned: If you take a ward, keep it---or find a worthy keeper.}
\tenet{4}{Debt Is Not Silence: Confess, repair, and let the record show it.}
\tenet{5}{No Erasures: Do not profit by deleting names, histories, or lineages.}

\subsection{Runekeeper: What You Get}

When you swear to the Penitent Lich as Runekeeper, you learn to inscribe small, portable bindings called runes. Each rune is a single-use inscription prepared during downtime or a quiet scene.
\begin{itemize}
    \item \textbf{Preparing Runes:} In downtime, prepare 2 runes from your list; carry at most 3 at a time. You may overwrite an unspent rune by marking Obligation to the Patron.
    \item \textbf{Invoking Runes:} Invoking a rune is usually Dominant and takes a quick action unless the scene is in chaos (then Dominant).
    \item \textbf{Interacting with Position/DV:} Runes usually raise Position one step or give +1 effect towards their stated sphere.
\end{itemize}

\subsection{Devotions \& Boons}

Track your standing with the Penitent Lich as Favor [0--5]. Increase Favor by naming the harmed, choosing restitution, keeping hard watches, or completing a rite of remembrance. Decrease Favor by erasing names, refusing confession, or striking down a pleading foe without terms.

\devotion{0}{Initiate of the Ledger}
\rite{Patron Gift --- Rune: Mark of Witness}{Chalk or ash mark that steadies Position once per scene when you defend the powerless or truth.}
\rite{Prayer---Soft Candle}{Once per session, dim a scene's hostility: -1 segment from an active Doom/Alert/Ward sub-clock after you speak who is being protected.}

\devotion{1}{Keeper of the Door}
\rite{Rune: Threshold}{Place on a doorway/line. For one scene, allies crossing it get +1d to resist separation or ambush.}
\rite{Boon---Confessor's Ears}{In a social exchange where you admit a fault, gain +1d to read motive (Insight-type handles) this scene.}

\devotion{2}{Binder of Names}
\rite{Rune: True Name}{Inscribe a true name on cloth/metal. For one action that targets the named (aid or oppose), choose: +1 effect or steady Position. If used to harm, mark Stain unless you offered terms first.}
\rite{Boon---Ledger Glimpse}{Once per session, ask the GM: ``What would restitution look like here?'' Gain a concrete, actionable step.}

\devotion{3}{Warden of Restitution}
\rite{Rune: Redress}{Break this seal to clear 2 segments from a Debt/Obligation clock you owe---but you must immediately perform a small, public act of repair.}
\rite{Boon---Vigil}{When you keep watch over a warded thing for a full scene, everyone on your side begins the next scene one Position higher.}

\devotion{4}{Litany-Bearer}
\rite{Rune: Litany}{Speak a short list of rightful names. For the next minute, your side treats the scene's first Desperate roll as Dominant. Consumes one distinct name from your Name Ledger.}
\rite{Boon---Mercy Carried Forward}{When you choose a nonviolent exit in a grievous scene, remove 1 segment from a Memory/Alert clock.}

\devotion{5}{Keeper of the Last Light}
\rite{Rite: The Thirty-Spoken}{If your Ledger holds 30 true names from a single penitential place, you may resolve a confrontation with persuasion instead of violence; on success, also refresh one spent rune.}
\rite{Boon---Absolution's Edge}{Once per arc, declare immunity to fear/charm for your party in a scene where you speak for the harmed.}

\subsection{Runes (Runekeeper List)}

Pick from these when you prepare:
\begin{itemize}
    \item Witness: Mark a surface; your side gets +1 effect to protect witnesses or evidence this scene.
    \item Threshold: See Favor 1.
    \item True Name: See Favor 2.
    \item Redress: See Favor 3.
    \item Peace: Anyone who voluntarily sheaths a weapon within arm's reach of this rune gains +1d to their next social action in this scene.
    \item Rest: Set by a body or grave; undead entering must test (DV 2). On fail, they hesitate and reveal who they were.
    \item Atonement: Place on an object tied to harm; returning it to its rightful keeper steadies Position for your group and creates a minor Boons token.
\end{itemize}

\subsection{Rites \& Invocations}
\rite{Candle of Remembrance (scene)}{Light a small candle and speak a harmed name. For this scene, you may re-roll 1s once on a social or investigation action that centers that person.}
\rite{Seal of Safe Passage (scene)}{Trace a circle and recite three names you have helped. For the next crossing (gate, bridge, alley), reduce the first Ward/Alert sub-clock you trigger by 1.}
\rite{Confessor's Vigil (downtime)}{Keep watch at a threshold named in your Ledger. Clear 1 stress/strain and prepare 1 extra rune next downtime.}

\subsection{Compels \& Taboos}
\compel{Offer a boon or clear 1 segment of an impending trouble if the Runekeeper attempts a nonviolent option first. Refusing marks Stain or loses 1 Favor (player's choice).}
\taboo{Erase a name for gain; profit from falsified records; abandon a sworn watch without successor; kill a pleading foe without terms.}

\subsection{Offerings \& Obligations}
\offering{Restore a headstone or ledger page.}
\offering{Return a stolen token (ring, book, oath-stone).}
\offering{Keep a night's watch for someone who cannot.}
\offering{Publicly confess a relevant fault and make a concrete repair.}

\obligation{Mark Obligation [4] when you take aid without an offering. Fulfilling a listed offering clears 2 segments.}

\synergy{If your party earned the Glaive of the Thirty, you may speak a Ledger name while guarding or striking to gain +1 effect or steady Position (once/scene). Re-using the same name before the next arc marks Stain.}

\subsection{Using This Patron in the Lich's Dungeon}
\begin{itemize}
    \item \textbf{Name Ledger:} Keep three columns: Harmed / Helped / Unresolved. Moving names from Unresolved → Helped is how you organically earn Favor.
    \item \textbf{SB Spending Guidance:} Spend SB to reveal truths, tighten oaths, or summon witnesses, not just to spike harm.
    \item \textbf{Nonviolent First:} With this Patron, players should feel tangible benefits for choosing mercy---Position bump, DV hints, and clock relief.
\end{itemize}

\subsection{Example Invocations}
\begin{itemize}
    \item ``By the breath that warmed ALYSSA WHITE-LILY's ring, let no vow be mocked while this mark holds.''
    \item ``Threshold, keep---the harmed may pass before the hungry.''
    \item ``Ledger, open---show me what restitution looks like, not what is easy.''
\end{itemize}

\subsection{GM Notes (Quick)}
\begin{itemize}
    \item Favor should move. Reward visible repair. Make taboos costly but recoverable through clean acts.
    \item When in doubt, let a rune raise Position or grant +1 effect toward mercy, truth, or watchkeeping rather than raw damage.
\end{itemize}

% Penitential Corruption
\section{Penitential Corruption --- Tables \& Rules}

Use this when the dungeon's necromancy stains the soul, when names are erased, or when the party spurns mercy. It integrates with Position, DV, SB, the Deck of Consequences, and the Name Ledger.

\corruptiontitle{How Corruption Works}
\begin{itemize}
    \item \textbf{When to Roll:} Gain Minor Corruption on small cruelties (breaking etiquette of remembrance, exploiting grief, petty desecrations). Gain Major Corruption for taboos (erasing a name, killing a pleading foe without terms, shattering a ward, stealing a ritual for profit, phylactery desecration without rites).
    \item \textbf{Track:} Each PC keeps a Corruption [0--6] track. Each roll adds +1 Corruption first, then you roll on the relevant table. Some results add more.
    \item \textbf{Stain vs. Corruption:} Stain is a narrative tag that can clear through play; Corruption is a mechanical track with thresholds below.
    \item \textbf{GM SB Hook:} In scenes of cruelty or erasure, the GM may spend 2 SB to force a Minor Corruption roll (once/scene).
    \item \textbf{Name Ledger Hook:} At Corruption ≥3, you cannot move any name from Unresolved → Helped without public restitution.
\end{itemize}

\threshold{3}{Choose a Mark (visible stigma). Deception about your past suffers DV +1. You cannot exceed Favor 3 with the Penitent Lich until you reduce to ≤2.}
\threshold{6}{You are Oath-Bound Revenant until cleansing (see below): all violent actions start Desperate; Glaive of the Thirty (if held) goes dormant; a personal Haunt follows you as a GM tool.}

\textbf{Cleansing Corruption (pick/stack as fiction allows)}
\begin{itemize}
    \item Public Confession + Concrete Restitution: -1
    \item Bury the Unburied / Restore a Headstone or Ledger Page: -1
    \item Keep a Named Vigil (one full scene) at a relevant threshold: -1
    \item Return a Stolen Token to its rightful keeper: -1
    \item Restore a Lineage / Undo an Erasure: -2
    \item The Thirty-Spoken (leader only): set Corruption to 0
\end{itemize}

\corruptiontitle{Table A --- Minor Corruption (d12)}
Roll when you take a petty cruelty or minor desecration.
\begin{enumerate}
    \item Cold Breath: Your breath steams in warm rooms. The first time you draw a weapon each scene, tick Alert +1 segment.
    \item Ink-Fingers: Black ink stains your hands. Evidence-type actions suffer DV +1 unless you start by admitting a relevant fault.
    \item Wrong-Whisper: Echoes misname you. Sway vs any harmed party is -1d unless you begin with acknowledgment; on acknowledgment, steady Position.
    \item Guttering Candle: Nearby flames flutter toward you. When you lie about the past, immediately draw the Deck of Consequences.
    \item Numb Grip: Fingers won't warm. Your first delicate action each scene (Tinker/Theft/etc.) starts Dominant unless you carry a token of the harmed.
    \item Cold Ring: A band bites your finger. If you break a promise this scene, tick a Memory Collapse [4] sub-clock +1.
    \item Skittering Shadow: Your shadow jitters. Stealth starts Dominant unless you spoke a true name today.
    \item Frosted Mirror: Your reflection fogs first. When acting from self-deception, DV +1; if you speak your true name, ignore this for the scene.
    \item Ash in Pocket: Coin spent for selfish gain becomes an IOU; tick Obligation +1 segment (to anyone harmed by the scene).
    \item Blurred Margins: Names blur at the edges of pages. Once/scene the GM may ask, ``Whose name is missing?'' If you cannot answer, Position -1.
    \item Hollow Step: Floors creak beneath you. Endure-type resists cost +1 stress/strain unless defending the powerless.
    \item Bleeding Ledger: When you destroy a record, mark Stain and roll Minor Corruption again.
\end{enumerate}

\corruptiontitle{Table B --- Major Corruption (d12)}
Roll when you violate a taboo or commit a grievous erasure.
\begin{enumerate}
    \item Counting Shadow: Your shadow counts softly. At Dusk, accept a Compel to attempt a nonviolent option first or mark Stain.
    \item The Unlit Candle: A thirty-first candle won't catch. Undead prefer you as a target, but you gain +1 effect on Rites after confession.
    \item Tongue of Ash: Mispronouncing a true name adds +1 segment to Persuasion/Memory Collapse and locks that name until restitution.
    \item Name-Scar: Healed wounds etch a name you can't say. Choose an NPC's name you are unable to speak until you perform a fitting repair.
    \item Frosted Vows: Your touch chills vows. Allies start the next social scene Dominant unless you consecrate with Rune: Witness.
    \item Mirror-Shard Soul: On being reduced to 0, you split into two conflicting drives for one scene (GM tool) until healed by a Vigil.
    \item Debtor's Sigil: Mark Obligation [4] to the Patron. Until cleared, you cannot invoke the same Litany or Glaive name twice per arc.
    \item Name-Eater: On a critical failure, one random name moves from Helped → Unresolved; gain +1d to your immediate follow-up and mark Stain.
    \item Ash Crown: Authority senses your usurpations: DV +1 vs authority until you abdicate something tangible.
    \item Counting Bells: Time slips. Each long action you take also ticks Alert/Ward +1 unless a harmed witness is present.
    \item Empty Pall: Your healing helps half unless accompanied by a spoken apology tied to the scene.
    \item Candle Tax: SB spent against you in a memory-scene has +1 potency (or GM pays 1 less SB) unless a PC reads a harmed name aloud at scene start.
\end{enumerate}

\corruptiontitle{Party Miasma (Optional Group Track)}
When multiple PCs commit erasures in one episode, tick Miasma [0--6].
\begin{itemize}
    \item \textbf{3+:} All scenes start Position -1 until the group performs a public act of repair together.
    \item \textbf{6:} The dungeon rejects the party---advance each active clock +1 and introduce a roaming Haunt until you restore a lineage or bury the unburied.
\end{itemize}

\corruptiontitle{Deck Integration (Quick)}
When a result calls for a Deck draw, interpret: Hearts---remembrance, Clubs---pressures, Diamonds---obligations, Spades---omens/necromancy. Use rank for severity.

\corruptiontitle{Runekeeper \& Glaive Interactions}
\begin{itemize}
    \item Runekeeper: Rune: True Name and Rune: Litany cannot be used to harm without offering terms; doing so adds +1 Corruption (then roll Minor).
    \item Glaive of the Thirty: For each point of Corruption, remove one distinct name from being used with Litany Edge until the next arc or until cleansed.
\end{itemize}

\corruptiontitle{GM Usage Notes}
\begin{itemize}
    \item Point to visible marks; make Corruption felt in mirrors, candles, ledgers, and vows.
    \item Make cleansing specific and public; reductions should be stories, not toggles.
    \item Keep the table punchy: roll once, narrate hard, show how mercy re-opens safe play.
\end{itemize}

% Utaran Empire Brief
\section{Utaran Empire --- GM Brief (Canon: 300 Years Ago)}

\textbf{Canon Snapshot}
\begin{itemize}
    \item \textbf{Utaran Empire}: A cosmopolitan river-ringed imperium of marble forums, oath-courts, civic cults, and fortified harbors. After centuries of overreach and factional intrigue, it \textbf{collapsed ~300 years ago} through \textbf{plague}, \textbf{sieges}, and \textbf{political purges}. No single event ended Utaran---its fall was a long unmaking.
    \item \textbf{Aftermath}: Provincial secessions; governors declaring sovereignty; temples shuttered; civic registries burned or rewritten; families fleeing along the river roads; mosaics chipped clean of disgraced names.
    \item \textbf{Why it matters here}: The Penitent Lich's worst decisions occur during this decline: the \textbf{Siege of Ecktoria}, unlawful executions, forbidden necromancy in the census vaults, and emergency edicts that cost lives. \textbf{The dungeon never incriminates directly}; it suggests guilt through echoes of administration, funerary rites, silenced names, and half-erased decrees.
\end{itemize}

\textbf{Visual \& Material Motifs (Late-Utaran)}
Use these to imply "Utaran" without naming it:
\begin{itemize}
    \item \textbf{River-Stone Masonry}: Gray-blue stone with silver mica; civic walls and courts always laid in \textbf{triple courses}.
    \item \textbf{Icon of the Noon Bell}: A bronze tower-bell whose chime is famously solemn; inscriptions claim "Noon judges all."
    \item \textbf{Laurel-Circle Crest}: A civic seal of interlocked laurels and a central keyhole---often \textbf{cracked cleanly by heat}.
    \item \textbf{Tri-Mark Script}: Formal bureaucratic hand marked by three flicks at line-ends; common on decrees, censuses, and military oaths.
    \item \textbf{Ash-Snow}: Pale ash drifting like snowfall; warms to the scent of river clay. Implies burning archives, cremated dead, and censored history.
    \item \textbf{Unfinished Mosaics}: Walls where tesserae show immaculate craftsmanship, then abruptly stop---budgets frozen, artisans vanished.
\end{itemize}
\textit{Fold these motifs into Anchors, Senses, and Name-as-Key clues. Mercy or restoration makes patterns more legible.}

\textbf{Dating Clues (300 Years Ago)}
\begin{itemize}
    \item \textbf{Coinage}: Greened bronze \textit{minas} stamped with the Noon Bell; edge-marks filed into threes.
    \item \textbf{Ledgers}: Oak-gall ink that "brown-ghosts" through thin civic vellum.
    \item \textbf{Stonework}: Every \textbf{tenth block scored thrice}---a construction code of late-Utaran engineers.
    \item \textbf{Edicts}: Margins framed by \textbf{Tri-Mark} calligraphy; censored names scraped thin.
\end{itemize}

\textbf{Mechanical Hooks}
\begin{itemize}
    \item \textbf{Imperial Research}: When a PC frames an action as \textit{late-Utaran scholarship} and presents a relevant artifact (coin, seal-fragment, ledger scrap, river-stone chip), grant \textbf{DV −1} or \textbf{Position +1} for Insight/Evidence once per floor.
    \item \textbf{Mercy → Clarity}: If a scene resolves via \textbf{mercy, restraint, or civic restoration}, reveal a clearer Utaranan clue (correct title, intact seal, a name no longer scratched away).
    \item \textbf{Utaran Tokens}: Fragments of laurel-seals, filings from Noon Bell bronze, bits of scorched ledger. \textbf{3 tokens} may be spent for \textbf{+1 progress} on the Final Persuasion clock, or to ask: \textit{"What restitution would honor the fallen city?"}
\end{itemize}

\textbf{Language \& Titles (for Name-as-Key)}
Scatter these into inscriptions and NPC recollections---never say "Utaran" out loud:
\begin{itemize}
    \item \textbf{Offices}: \textit{Praetor of the Laurel-Circle; Warden of the River Ring; Master of the Imperial Archives; Censor of Noon}.
    \item \textbf{Civic Phrases}: \textit{"By ring and river," "Noon judges all," "Three marks bind truth," "In service to the Circle."}
\end{itemize}

\textbf{GM Guidance}
\begin{itemize}
    \item \textbf{Imply, Never Accuse}: No NPC says the Lich destroyed the Empire. The players infer it through \textit{abandoned tribunals, burnt ledgers, missing names, unfinished mosaics}.
    \item \textbf{Echoes, Not Exposition}: Choose \textbf{2--3 motifs} (Noon Bell, river-stone, laurel seals) and repeat them across distant floors. Recognition is the payoff.
    \item \textbf{Restitution Scenes}: Let players mend civic wounds---re-engrave censored names, restore a bell-rope, finish a mosaic, return confiscated funerary seals. These small restorations make the fall visceral.
\end{itemize}

\textbf{Tie-ins}
\begin{itemize}
    \item \textbf{Runekeeper Patron}: +1 Favor when a PC restores a civic record \textbf{without excising shameful entries}. History is honored, not rewritten.
    \item \textbf{Corruption}: If a PC \textbf{uses Utaranan clues} but refuses restitution or erases names, trigger \textbf{Ash-Crown}-style Major Corruption.
    \item \textbf{Final Floor}: The 30 candles form a \textbf{ring-wall mosaic} around an unlit phylactery set on river-stone. During the \textbf{Thirty-Spoken}, the \textbf{Bell of Noon} finally chimes clean---its tone unbroken for the first time.
\end{itemize}

% Running Notes & Conversions
\section{Running Notes \& Conversions}
\subsection{Scaling Clocks}
If the group is breezing through, upgrade [4] → [5] or add a Ward/Alert sub-clock keyed to a floor's trap.

\subsection{Position Drift}
Mercy, confession, and restitution raise Position next floor; cruelty or evasion drop it. Keep a simple toggle per PC.

\subsection{SB Spend Menu (Dungeon Flavor)}
\begin{description}[leftmargin=*, style=unboxed]
    \sbitem{1}{Memory whispers; +1 Supply segment; evidence goes missing.}
    \sbitem{2}{Alarm raised; lose cover; a lesser guardian arrives; vow-echo asks a costly question.}
    \sbitem{3}{Reinforcements; key gear breaks; rail tick to a larger tragedy.}
    \sbitem{4+}{Trap springs; authority/omen arrives; scene shifts to harsher truth.}
\end{description}

\subsection{Boons as Compassion}
When players console, bury, confess, or repair---hand out Boons. Let them convert Boons to stabilize Position at moral cruxes.

\subsection{Deck Cadence}
Aim for 1--2 Deck draws per session unless the table wants chaos. Translate results into fiction: omens, debts, vows---not just penalties.

% Prep Shortcuts
\section{Prep Shortcuts}
\subsection{Make a Name Ledger}
Three columns---Harmed, Helped, Unresolved. Move names between columns as choices land.

\subsection{Keep a Shard Kit}
Mirror shard, ring, ledger scrap, bone-petal---physical tokens that move between floors to track themes.

\subsection{Sketch the lich in three moods}
Proud, Weary, Resolved. Check one at the end of each session; bring that mask to Floor 30.

% Rewards & Advancement
\section{Rewards \& Advancement}
This adventure is a major arc: award arc XP at resolution. Each floor can also grant small, flavorful boons (temporary tags, one-use rites, favors). Compassionate resolutions often trade immediate safety for long-term strength---lean into it.

% Appendix
\section{Appendix: Quick Floor Hooks by Pillar}
\begin{description}[leftmargin=*, labelwidth=4cm, style=standard]
    \item[Combat-forward:] 3, 10, 12, 15, 17, 28
    \item[Social-forward:] 2, 4, 7, 13, 18, 30
    \item[Puzzle/Investigation:] 1, 5, 8, 11, 19, 29
\end{description}

Use as a one-shot (pick 6--8 floors), a short arc (12--16 floors), or the full penitence (all 30). The lich isn't "defeated"---he's convinced.

\end{document}

