\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{tcolorbox}
\hypersetup{colorlinks=true, linkcolor=black, urlcolor=blue}

\newcommand{\Part}[1]{\section*{#1}}
\newcommand{\Clock}[2]{\textbf{#1} [#2]}
\newcommand{\Term}[1]{\textbf{#1}}
\newcommand{\DV}{\Term{DV}}
\newcommand{\Position}{\Term{Position}}
\newcommand{\SB}{\Term{SB}}
\newcommand{\Diamond}{\Term{Diamond}}
\newcommand{\Tier}{\Term{Tier}}

\definecolor{softgray}{RGB}{245,245,245}
\definecolor{inkblue}{RGB}{20,40,120}
\definecolor{blood}{RGB}{130,10,10}
\definecolor{forest}{RGB}{16,100,60}

\tcbset{colback=softgray, colframe=black, boxrule=0.4pt, arc=2mm}

\begin{document}

\begin{center}
{\LARGE \textbf{The Silk \& Velvet Arc}}\\[4pt]
{\large An Urban Intrigue Adventure for \emph{Fate's Edge}}\\[2pt]
\textit{Tiers I--III \;|\; 6--10 sessions \;|\; Social maneuvering, ritual peril, political theatre}
\end{center}

\hrule
\vspace{0.5em}

\subsection*{What This Is}
A complete, table-ready arc set across the trade-bridge city of \textbf{Silkstrand}, a night of high etiquette at \textbf{Ravenswood Manor}, and the stair-ruled metropolis of \textbf{Thepyrgos}. Themes: contracts and crowds; masks and glamour; law as performance. The adventure emphasizes \Position{} play, interlocking \Term{clocks}, and a flavorful \SB{} economy.

\subsection*{At a Glance}
\begin{itemize}
  \item \textbf{Part I --- A Song of Poison and Plague} (Tier I--II): Urban investigation as the Gemcutters and Alchemists edge toward riots while a bard of Desire plays both sides.
  \item \textbf{Part II --- The Gilded Cage} (Tier II--III): Survive a night of velvet hospitality; each room is a social encounter with taboos and tells.
  \item \textbf{Part III --- The Crimson Masque} (Tier III): A public ritual threatens to swap the souls of the city's elite with the troupe; stop it on stage before the final bell.
\end{itemize}

\begin{tcolorbox}
\textbf{System Notes.} Use the game's Position ladder and \DV{} to make leverage matter. City-specific \Diamond{}s represent codified outcomes (permits, seats, escorts) that shift \Position{} without a roll. Clocks pace consequences. The GM spends \SB{} when 1s appear to twist crowds, shift law, or advance rails.
\end{tcolorbox}

\subsection*{Adventure Background}
Silkstrand sits over dyed waterways and legal bridges; permits and seals are visible power. Two guilds---\textbf{Gemcutters} and \textbf{Alchemists}---blame each other for ``poisoned'' water. \textbf{Sel Virelle}, a charismatic devotee of Desire, orchestrates paranoia to ripen the city for a prestige heist by the \textbf{Crimson Masque} troupe, agents of Glamour. Their endgame plays out in Thepyrgos, where bell-law and stair-rights can crown or crush reputations overnight.

\subsection*{Session Count \& Tier Flow}
Intended for 6--10 sessions. Start at Tier I, crest at Tier III. Each part includes typical \DV{} and \Position{} guidance to ease scaling.

\section*{Core Tools for the GM}

\subsection*{Position \& DV Quick Reference}
\vspace{-0.5em}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Position} & \textbf{Typical DV} & \textbf{Examples of Fictional Advantage}\\
\midrule
Dominant & \(\Tier+2\) & You hold permits, the crowd favors you, law is on your side.\\
Controlled & \(\Tier+3\) & Proper introductions, stable footing, watch is neutral.\\
Risky & \(\Tier+4\) & Time pressure, hostile witnesses, partial leverage.\\
Desperate & \(\Tier+5\) or more & No permits, hostile crowd, taboo broken, the clock is high.\\
\bottomrule
\end{tabularx}

\subsection*{Typical DVs by Part}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Part} & \textbf{Common DV Range} & \textbf{Notes}\\
\midrule
I: Silkstrand & 3--4 & Documents and Diamonds can drop DV by 1 before rolling.\\
II: Ravenswood & 4--5 & Etiquette taboos raise DV; reading tells can lower DV by 1.\\
III: Thepyrgos & 5--6 & Public ritual actions start high; dismantling pillars lowers DV.\\
\bottomrule
\end{tabularx}

\subsection*{Diamonds as Leverage}
\begin{tabularx}{\linewidth}{@{}l X l@{}}
\toprule
\textbf{Diamond} & \textbf{Fictional Authority} & \textbf{Effect}\\
\midrule
Matron's Audience & You may petition Silkstrand's Matron in session. & Shift \Position{} one step up once/scene; \(-1\) DV on petitions.\\
Arsenal Key & Authorized access to a guild armory or lockup. & Once/session, declare gear or evidence ``on hand''.\\
Bellmark & A stamped right to invoke bell-law in Thepyrgos. & Once/scene, freeze a crowd or call watch for 1 beat.\\
Stair-Pass & Priority use of a named stair during events. & Downshift \Position{} consequences from crowd pressure.\\
Escort Writ & Legal escort through markets/bridges. & Ignore one blockade or mob once/session.\\
\bottomrule
\end{tabularx}

\subsection*{SB Spend Menus (Use When 1s Appear)}
\noindent\textbf{Silkstrand}
\begin{itemize}
\item Bridge duel erupts nearby; escalate a scene or split the party.
\item Market panic closes a route; add 1 to a public clock.
\item Dyewater omen flips the crowd's mood against the PCs.
\item Notary ``corrects'' a permit; \Position{} downshift for one exchange.
\end{itemize}
\noindent\textbf{Ravenswood Manor}
\begin{itemize}
\item Etiquette whisper spreads; tick \Clock{Etiquette Breach}{1}.
\item Hospitality favors a rival; your next roll is at higher DV.
\item Servant intervention reroutes you to a new room at a cost.
\end{itemize}
\noindent\textbf{Thepyrgos}
\begin{itemize}
\item Bell strikes early; advance the Act clock.
\item Stair-right revoked temporarily; block or split routes.
\item Law change mid-scene; transform a planned action's \Position{}.
\end{itemize}

\subsection*{Clocks to Start With}
Keep clocks lean at first; layer more as play demands.
\begin{itemize}
\item \Clock{City Unrest}{6} (Silkstrand)
\item \Clock{Poisonings}{4} (Silkstrand)
\item \Clock{Sel's Glamour Net}{4}
\item Choose \emph{one} guild reputation to track first: \Clock{Gemcutters' Favor}{4} \emph{or} \Clock{Alchemists' Favor}{4}
\end{itemize}

\section*{Part I --- A Song of Poison and Plague (Tier I--II)}

\subsection*{Strong Hooks (Choose 1--2 to Open)}
\begin{itemize}
\item \textbf{Dyewater Omen:} The river runs a prophetic hue; the crowd demands action from anyone wearing steel or sigils.
\item \textbf{Bridge Duel Gone Wrong:} A fatal misstep during a public duel implicates a guild; the watch seeks neutral witnesses (the PCs).
\item \textbf{Missing Ledgers:} Archivolt records vanish; without them, permits are void and riots loom.
\end{itemize}

\subsection*{Core Clocks}
\Clock{City Unrest}{6}, \Clock{Poisonings}{4}, \Clock{Sel's Glamour Net}{4}, plus one Guild Favor clock.

\subsection*{Key NPCs}
\begin{itemize}
\item \textbf{Guildmistress Kavra (Gemcutters)} --- \emph{Desires:} stability, monopoly on luxury. \emph{Leverage:} bravos, patronage. \emph{Tell:} rubs thumb over a signet when lying.
\item \textbf{Master Orun (Alchemists)} --- \emph{Desires:} legitimacy, funding. \emph{Leverage:} inspectors, antidotes. \emph{Tell:} recites measurements under stress.
\item \textbf{Sel Virelle (Bard of Desire)} --- \emph{Desires:} acclaim, volatility. \emph{Leverage:} charmed witnesses. \emph{Tell:} mirrors phrasing to draw agreement.
\end{itemize}

\subsection*{Scenes}
\paragraph{Redwater Audit.} Enter the dyeworks with forged or earned permits. \Position{} starts Controlled; DV 3--4. On success with cost, allow entry but spend \SB{} to tick \Clock{Poisonings}{+1} or call an Alchemist observer. \emph{Magic:} Alchemical detection, illusions to mimic stamps, or subtle glamours to steer crowds shine here.

\paragraph{Alleyway Serenade.} Sel ``rescues'' a bystander from a mob to win PC trust. Reading her tells can downshift DV. Failure-forward: the PCs earn a \Diamond{} (Escort Writ) but tick \Clock{Sel's Glamour Net}{+1}.

\paragraph{Three-Queens Duel.} A formal bridge duel settles a permit dispute. Stakes: purse, permit, or pride. Victory can clear 1 from a Guild Favor clock; loss still reveals a planted second's testimony.

\paragraph{Archivolt Ledgers.} Track falsified entries to a clerks' circle. Options: bribe (Diamonds), debate in public (drop DV by 1 with a crowd), or stealth (risk Desperate). On a miss, the ledgers you \emph{do} find implicate an ally---use as leverage later.

\subsection*{Climax Options}
\begin{itemize}
\item \textbf{Public Debate at Lion Bridge:} Expose Sel. If you sway the crowd, cool \Clock{City Unrest}{-2}. If you fail, you still win a legal foothold but \Clock{Glamour Net}{+2}.
\item \textbf{Quiet Sabotage:} Collapse distribution of forged evidence; gain a permanent \Diamond{} (Matron's Audience) but sour one guild permanently.
\item \textbf{Arrest \& Bargain:} Trade Sel's testimony for guild peace; start Part II with social capital at Ravenswood.
\end{itemize}

\begin{tcolorbox}
\textbf{Fail-Forward Examples (Part I).}
\begin{itemize}
\item \emph{Audit Miss:} You are barred, but a junior clerk slips you an \emph{appointment chit} (treat as a temporary \Diamond{})—and the watch starts tailing you.
\item \emph{Duel Miss:} You lose the bout, but the crowd turns on the victor over a rule infraction; shift \Position{} up in the next public scene.
\item \emph{Ledger Heist Miss:} You grab the wrong ledger; it names a friend. Keep it and gain leverage, or return it and earn a favor.
\end{itemize}
\end{tcolorbox}

\section*{Part II --- The Gilded Cage (Tier II--III)}

\subsection*{Structure}
Track \Clock{Midnight Duel}{8} (the rival's plot) and \Clock{Etiquette Breach}{4} (your faux pas). Every room lists an \emph{Intent}, a \emph{Taboo}, and a \emph{Tell}. Violating taboos increases DV or ticks \Clock{Etiquette Breach}{+1}. Reading tells can reduce DV by 1.

\subsection*{Rooms}
\paragraph{Conservatory of Masks.} \emph{Intent:} choose alliances. \emph{Taboo:} wearing the same mask twice. \emph{Tell:} a rival never mirrors your stance. \emph{Magic:} Disguises and glamours gain a clear advantage; clever illusions can ``prove'' etiquette creatively.

\paragraph{Library of Additions.} \emph{Intent:} alter contracts with lawful flourishes. \emph{Taboo:} citing a text you cannot produce. \emph{Tell:} a steward coughs before lies. \emph{Magic:} Divination or memory magic can surface the missing clause; ritual ink resists forgery unless properly countered.

\paragraph{Ballroom of Vows.} \emph{Intent:} demonstrate poise. \emph{Taboo:} cutting a trio dance to two. \emph{Tell:} a partner flinches at touch. \emph{Magic:} Subtle enchantments risk taboos; protective magic can prevent social harm without force.

\subsection*{Outcomes}
\begin{itemize}
\item \textbf{Favor the Host:} Gain introductions---begin Part III with a Bellmark.
\item \textbf{Favor the Challenger (Glamour-touched):} Receive access to the Masque's rehearsal; start Act I behind the curtain.
\item \textbf{Seize the Venue:} Control the guest list that matters later; start with \Position{} at Dominant for Act I scene one.
\end{itemize}

\begin{tcolorbox}
\textbf{Fail-Forward Examples (Part II).}
\begin{itemize}
\item \emph{Mask Faux Pas:} \Clock{Etiquette Breach}{+1}, but you notice a hidden side-door used by rivals.
\item \emph{Contract Gaffe:} Your addition is rejected publicly; the steward, impressed by your boldness, offers a private reading later (lower DV once).
\item \emph{Dance Misstep:} You stumble, but hear a whispered cue that reveals the rival's schedule.
\end{itemize}
\end{tcolorbox}

\section*{Part III --- The Crimson Masque (Tier III)}

\subsection*{Act Structure}
Run as a three-day countdown to the final performance.
\begin{itemize}
\item \textbf{Act I (Day --3):} Infiltrate rehearsals, court patrons, secure stair access.
\item \textbf{Act II (Day --2/--1):} Decode the rite; seed counters in law, masks, or music.
\item \textbf{Act III (Performance Night):} On-stage skill challenge; illusions are real; dismantle ritual pillars.
\end{itemize}

\subsection*{Ritual Pillars (Break One to Stop the Rite)}
\begin{itemize}
\item \textbf{Bell Cadence:} The final peal synchronizes souls. Counter with bell-law, a forged bellmark, or a timed silence.
\item \textbf{Mask Index:} Masks hold targeted identities. Swap labels, smudge index entries, or sanctify a mask to render it immune.
\item \textbf{Applause Escrow:} The ovation stores consent. Split the crowd, mute applause, or redirect acclaim to the troupe alone.
\end{itemize}

\subsection*{On-Stage Challenge (Guidance)}
Start at DV 6 (Desperate). Each broken pillar lowers DV by 1 and shifts \Position{} up. Failure-forward should always change the stage: a mask cracks, a bell peals off-time, an actor falters.

\subsection*{Endings}
\begin{itemize}
\item \textbf{Expose \& Redirect:} Law halts the rite; only the willing swap. Reputation boons carry forward.
\item \textbf{Strike in Glamour:} Duel within unfolding illusions while allies sever a pillar.
\item \textbf{Take the Price:} Bargain with Glamour or Desire; end the rite, owe a favor.
\end{itemize}

\section*{NPC Roster (Quick-Use)}

\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Name} & \textbf{Desires, Leverage, Tell, Voice}\\
\midrule
Guildmistress Kavra & Stability; bravos, patronage; signet rub; clipped, economical speech.\\
Master Orun & Legitimacy; inspectors, antidotes; muttered measures; careful, exact.\\
Sel Virelle & Applause; charmed witnesses; mirrored phrasing; warm, rising cadence.\\
Lord/Host of Ravenswood & Prestige; hospitality law; fixed smiles; velvet, measured tone.\\
The Challenger & Territory; glamours, etiquette zealots; perfect posture; honeyed cruelty.\\
Masque Director & Immortality-by-swap; troupe fanatics; eye-contact tests; theatrical whisper.\\
\bottomrule
\end{tabularx}

\section*{Appendix A: Session 1 Outline (3 Hours)}

\begin{enumerate}
\item \textbf{Cold Open (15 min):} Choose a hook; paint the crowd; surface a visible Diamond (Escort Writ in a steward's hand).
\item \textbf{Investigation Beat (45 min):} Redwater Audit or Missing Ledgers. Offer Position shifts via Diamonds.
\item \textbf{Pressure Beat (30 min):} Alleyway Serenade or Bridge Duel challenge.
\item \textbf{Choice Beat (30 min):} Which guild gets aid? Start one Guild Favor clock.
\item \textbf{Climax (30--40 min):} Public mini-debate or duel; on any miss, spend \SB{} to tick \Clock{Sel's Glamour Net}{+1}.
\item \textbf{Fallout (10 min):} Hand out one Diamond; foreshadow Ravenswood with a sealed invitation.
\end{enumerate}

\section*{Appendix B: Printable Clocks}
\begin{tcolorbox}
\Clock{City Unrest}{6}\hfill\(\square\square\square\square\square\square\)\\[3pt]
\Clock{Poisonings}{4}\hfill\(\square\square\square\square\)\\[3pt]
\Clock{Sel's Glamour Net}{4}\hfill\(\square\square\square\square\)\\[3pt]
\Clock{Gemcutters' Favor}{4}\hfill\(\square\square\square\square\)\quad
\Clock{Alchemists' Favor}{4}\hfill\(\square\square\square\square\)\\[3pt]
\Clock{Midnight Duel}{8}\hfill\(\square\square\square\square\square\square\square\square\)\\[3pt]
\Clock{Etiquette Breach}{4}\hfill\(\square\square\square\square\)\\[3pt]
\Clock{Act Countdown}{6}\hfill\(\square\square\square\square\square\square\)
\end{tcolorbox}

\section*{Appendix C: Player-Facing Rumors (Silkstrand)}
\begin{itemize}
\item ``Gemcutters bribe the river with rubies; the water drinks poorly of greed.''
\item ``Alchemists dump antidote pre-cursors at dawn---no poison, just panic.''
\item ``A singer can still a riot if the key is true.''
\end{itemize}

\vfill
\begin{center}
\small{\emph{The Silk \& Velvet Arc} --- A Fate's Edge adventure.}
\end{center}

\end{document}