\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage{enumitem}
\hypersetup{colorlinks=true, linkcolor=black, urlcolor=blue}

\newcommand{\Part}[1]{\section*{#1}}
\newcommand{\Clock}[2]{\textbf{#1} [#2]}
\newcommand{\Term}[1]{\textbf{#1}}
\newcommand{\DV}{\Term{DV}}
\newcommand{\Position}{\Term{Position}}
\newcommand{\SB}{\Term{SB}}
\newcommand{\Diamond}{\Term{Diamond}}
\newcommand{\Tier}{\Term{Tier}}

\definecolor{softgray}{RGB}{245,245,245}
\definecolor{inkblue}{RGB}{20,40,120}
\definecolor{blood}{RGB}{130,10,10}
\definecolor{forest}{RGB}{16,100,60}

\tcbset{colback=softgray, colframe=black, boxrule=0.4pt, arc=2mm}

\begin{document}

\begin{center}
{\LARGE \textbf{Steps Into Sorrow}}\\[4pt]
{\large A Two-Part Exploration Arc for \emph{Fate's Edge}}\\[2pt]
\textit{Tiers II--IV \;|\; 6--9 sessions \;|\; Philosophy puzzles, planar hazards, rival expeditions}
\end{center}

\hrule
\vspace{0.5em}

\subsection*{What This Is}
A complete arc that begins on the legendary \textbf{Bridge of a Thousand Steps} and descends into the shadow-rift metropolis of \textbf{Seven-Sorrows}. It emphasizes \Position{} play, fail-forward travel, environmental puzzles, and escalating rival pressure. The finale centers on the \textbf{Orrery of Lost Worlds}.

\subsection*{At a Glance}
\begin{itemize}
  \item \textbf{Part I --- The Bridge of a Thousand Steps} (Tier II): Each step poses a moral riddle or personal reckoning; failure costs gear, memory, or time instead of lives.
  \item \textbf{Part II --- The Sunken City of Seven-Sorrows} (Tier III--IV): A thematic ``district crawl'' where sorrow-aspect challenges unlock the Orrery.
\end{itemize}

\begin{tcolorbox}
\textbf{System Notes.} \Position{} sets your approach to hazards and social scenes; treat \Diamond{}s as codified outcomes (keys, escorts, permits) that re-frame \Position{} before rolling. Clocks pace rivals, environment, and dread. When 1s appear, spend \SB{} to twist the route, split the party, or advance a rival \emph{rail}. Keep dice scarce by rewarding preparation.
\end{tcolorbox}

\section*{Core Tools for the GM}

\subsection*{Position \& DV Quick Reference}
\vspace{-0.5em}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Position} & \textbf{Typical DV} & \textbf{Examples of Fictional Advantage}\\
\midrule
Dominant & \(\Tier+2\) & Correct answer in hand; guide spirits favor you; rival stalled.\\
Controlled & \(\Tier+3\) & Stable footing; time to think; relevant lore or tokens.\\
Risky & \(\Tier+4\) & Clock pressure; partial clue; split attention.\\
Desperate & \(\Tier+5\) or more & Rift surge; taboo breached; rival in your face.\\
\bottomrule
\end{tabularx}

\subsection*{Suggested DV Ranges by Part}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Part} & \textbf{Common DV Range} & \textbf{Notes}\\
\midrule
I: Bridge & 3--5 & Late steps escalate; spend Diamonds to reduce DV for one team member.\\
II: Seven-Sorrows & 4--6 & Districts start at DV 4; the Orrery begins at DV 6, dropping per solved sorrow.\\
\bottomrule
\end{tabularx}

\subsection*{SB Spend Menus}
\noindent\textbf{Bridge of a Thousand Steps}
\begin{itemize}[leftmargin=*]
\item A step misinterprets intent; reroll at higher DV or pay a \emph{Toll} (gear lost, memory tithe, or Dread +1).
\item A crosswind of whispers moves a stone; party splits unless someone anchors with a confession.
\item Another traveler manifests (echo of a PC's past) and demands to ``walk with you'' as a complication.
\end{itemize}
\noindent\textbf{Seven-Sorrows}
\begin{itemize}[leftmargin=*]
\item Sorrow-echo duplicates a PC's last mistake in public, worsening crowd \Position{}.
\item Rival expeditions get ahead; tick \Clock{Rival Lead}{+1} and close a route.
\item Rift surge changes district aspect; DV +1 until checked by a rite or offering.
\end{itemize}

\subsection*{Rivals (Use 1--2)}
\begin{itemize}
\item \textbf{Black Ferrule Company} --- mercenary treasure-rights. \emph{Edge:} anchors, explosives. \emph{Tell:} iron ferrules on boots.
\item \textbf{Order of the Last Dawn} --- apocalyptic zeal. \emph{Edge:} sacramental wards. \emph{Tell:} ash-smudged fingers.
\item \textbf{Gulls of the Chart} --- scholar-cartographers. \emph{Edge:} partial maps, clever shortcuts. \emph{Tell:} salt-ink stains.
\end{itemize}
Create a \Clock{Rival Progress}{6} and a \Clock{Rival Lead}{3}. Each SB you spend can tick one; PCs can reduce them with victories or bribes.

\section*{Part I --- The Bridge of a Thousand Steps (Tier II)}

\subsection*{Frame}
The bridge is not continuous: each stone appears when a riddle is answered \emph{in deed or word}. Steps draw questions from the travelers' regrets and values. Failing a step means cost, not death.

\subsection*{Crossing Clock}
Use a \Clock{Crossing}{8} to track fatigue and conditions. Each setback ticks it. On fill, an enforced \emph{camp-on-air} scene occurs: safety at the price of a Toll.

\subsection*{Approaches}
Each step can be met by different approaches; match to playbooks and talents.
\begin{itemize}
\item \emph{Lore/Insight:} reason, recall, pattern. 
\item \emph{Resolve/Spirit:} confession, vow, acceptance.
\item \emph{Athletics/Wits:} leap, counter-gust, lateral thinking.
\item \emph{Magic/Rites:} sanctify a step, bind a doubt, mirror a truth.
\end{itemize}

\subsection*{Step Table (Roll or Choose)}
\begin{tabularx}{\linewidth}{@{}l X X@{}}
\toprule
\textbf{d6} & \textbf{Prompt} & \textbf{Consequence (on Cost/Miss)}\\
\midrule
1 & ``Which promise still binds you when none watch?'' & Lose a minor item you brought to impress; it sinks into the air.\\
2 & ``What price would you pay to know the end?'' & Memory tithe: choose a pleasant memory; it's hazy until journey's end.\\
3 & ``Whom did you fail to bury?'' & A sorrow-echo follows, repeating a last conversation at bad moments.\\
4 & ``What tool do you reach for too quickly?'' & Tools feel heavier; add +1 DV to \emph{force} approaches until confessed.\\
5 & ``What do you refuse to forgive in yourself?'' & Dread +1. A later apology to a party member clears it.\\
6 & ``What would you save if only one thing?'' & A stranger ahead now holds it; bargain or race (tick \Clock{Rival Lead}{+1}).\\
\bottomrule
\end{tabularx}

\subsection*{Signature Scenes}
\paragraph{The Philosopher's Truss.} Interlocked stones demand three truths told in sequence: self, companion, stranger. Ordering them well grants a \Diamond{} (Guide's Favor) that downshifts \Position{} once/scene on the bridge.

\paragraph{The Toll-Taker.} An old keeper asks for coin, confession, or a scar. Paying grants \emph{Anchored Footing} (ignore one split/shift SB). Cheating the Toll-Taker ticks \Clock{Crossing}{+2}.

\paragraph{Mirrored Step.} Each PC must answer the other's prompt. On a miss, swap burdens for a scene; the wrong person suffers the cost---often revelatory, never lethal.

\subsection*{Bridge Climax Options}
\begin{itemize}
\item \textbf{Leap of Fellowship:} Combine efforts (group action) to bridge a missing span. On cost, one PC leaves something behind (gear, memory, or favor).
\item \textbf{Answer in Deed:} Resolve a party member's step by acting it out now. Grants a group \Diamond{} (Shared Vow) usable once per part.
\item \textbf{Summon a Way:} Magic shapes a temporary stone. It lasts exactly as long as the caster holds their breath in honesty.
\end{itemize}

\section*{Part II --- The Sunken City of Seven-Sorrows (Tier III--IV)}

\subsection*{Frame}
Seven districts embody seven kinds of sorrow. Each solved sorrow weakens the rift and lowers the Orrery's DV by 1 (to a minimum of 3). Districts invite non-combat problem-solving; violence tends to multiply echoes.

\subsection*{City Clocks}
\begin{itemize}
\item \Clock{Rift Instability}{6} (surges close or rotate streets)
\item \Clock{Rival Progress}{6} (shared with Part I)
\item \Clock{Civic Memory}{7} (how much of the city remembers itself)
\end{itemize}

\subsection*{District Menu (Choose 3--5)}
\paragraph{District of Broken Promises.} Doorways recite oaths. \emph{Challenge:} correctly amend an oath to include true limits. \emph{SB:} doors bargain with rivals. \emph{Reward:} \Diamond{} (Right of Passage) through any one sealed portal.

\paragraph{Quarter of Final Regrets.} Plazas show last choices. \emph{Challenge:} stage a reenactment with a different ending that \emph{could} have been. \emph{Fail-forward:} the crowd insists on an encore at the worst time.

\paragraph{Hall of Unfinished Songs.} Melodies lack final lines. \emph{Challenge:} complete a refrain that heals rather than conquers. \emph{Magic:} resonant spells can take the melody's shape. \emph{Reward:} reduce \Clock{Rift Instability}{-2}.

\paragraph{Market of Unsent Letters.} Stalls sell words never mailed. \emph{Challenge:} purchase the right unsent letter and deliver it to a living recipient. \emph{Cost:} you must leave an unsent letter of your own. \emph{Reward:} \Diamond{} (Bearer's Seal).

\paragraph{Gardens of No Farewell.} Plants grow from farewells withheld. \emph{Challenge:} prune without killing: a careful \Position{} test. \emph{SB:} roots entangle gear; lose something symbolic.

\paragraph{Tower of Abandoned Vows.} Steps climb but do not descend; each level asks a vow renewed or released. \emph{Challenge:} name what continues, and what ends. \emph{Reward:} \Diamond{} (Clear Conscience): once, ignore a social taboo's penalty.

\paragraph{Drowned Forum.} Debate pits beneath shallow black water. \emph{Challenge:} argue submerged; breath sets the tempo. \emph{Reward:} set crowd mood for next public scene; reduce \DV{} by 1 for a debate.

\subsection*{The Orrery of Lost Worlds (Finale)}
A chamber of spheres, rings, and chains, orbiting a blank star. Start at DV 6; each district solved reduces DV by 1. Three \emph{Pillars} hold the mechanism in fate:

\begin{itemize}
\item \textbf{Axis of Names.} The city remembers who it was. \emph{Counter:} place true names recovered from letters and songs.
\item \textbf{Weights of Debt.} Tallies of bargains hold the orbits. \emph{Counter:} cancel or forgive a meaningful debt on-screen.
\item \textbf{Gnomon of Return.} A shadow-mark sets the city's angle. \emph{Counter:} align with the Bridge's rhythm (Shared Vow, Guide's Favor, or a sacrificial pause).
\end{itemize}

\subsection*{Outcomes}
\begin{itemize}
\item \textbf{Seal the Rift:} The city stabilizes as a twilight district accessible by rite. Each PC chooses one sorrow they have \emph{truly} laid down; gain a lingering boon.
\item \textbf{Map the Paths:} The Orrery points to other lost places. Rivals depart rich; you earn routes and obligations.
\item \textbf{Offer to a Patron:} Trade the Orrery's motion for a boon. The city changes---and so does the world nearby.
\end{itemize}

\section*{NPC Roster (Quick-Use)}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Name} & \textbf{Desires, Leverage, Tell, Voice}\\
\midrule
Captain Brann of the Black Ferrule & Contracts; explosives, anchors; boots scrape stone; blunt, soldierly.\\
Sister Ysol of the Last Dawn & Purity; sacrament wards; ash-smudge ritual; fervent, precise.\\
Cartographer Pell of the Gulls & Knowledge; partial charts; hums while plotting; eager, scattered.\\
The Toll-Taker & Toll paid; rules of passage; weighs your words; wind-dry, patient.\\
The Conductor of Unfinished Songs & Completion; choirs of echoes; gestures a downbeat; resonant hush.\\
\bottomrule
\end{tabularx}

\appendix 

\section{Session 1 (Bridge) --- 3 Hours}
\begin{enumerate}
\item \textbf{Arrival (15 min):} Establish why each PC must cross. Show a visible Diamond (Guide's Favor hint).
\item \textbf{First Five Steps (45 min):} Mix prompts 1--3. Offer a fail-forward that grants a temporary boon with a cost.
\item \textbf{Pressure Beat (30 min):} Wind shift splits the line; group action to reunite or pay a Toll.
\item \textbf{Set-Piece (30--40 min):} Philosopher's Truss; earn Guide's Favor or tick \Clock{Crossing}{+2}.
\item \textbf{Camp-on-Air (10 min):} On fatigue, rest with a choice of coin, confession, or scar.
\end{enumerate}

\section{Printable Clocks}
\begin{tcolorbox}
\Clock{Crossing}{8}\hfill\(\square\square\square\square\square\square\square\square\)\\[3pt]
\Clock{Rival Progress}{6}\hfill\(\square\square\square\square\square\square\)\quad
\Clock{Rival Lead}{3}\hfill\(\square\square\square\)\\[3pt]
\Clock{Rift Instability}{6}\hfill\(\square\square\square\square\square\square\)\\[3pt]
\Clock{Civic Memory}{7}\hfill\(\square\square\square\square\square\square\square\)
\end{tcolorbox}

\section*{Appendix C: Player-Facing Handouts}
\paragraph{Crossing Rights.} ``The bridge admits those who answer in deed. Pay with what weighs you most.''
\paragraph{Sorrow Map (Text).} A circular diagram of seven districts; connect any three to weaken the Orrery.

\vfill
\begin{center}
\small{\emph{Steps Into Sorrow} --- A Fate's Edge adventure.}
\end{center}

\section{GM Enhancements Overview}
\label{app:enhancements}
\textbf{Included options:}
\begin{itemize}
\item \textbf{XP milestone awards} for character progression.
\item \textbf{Magic path specializations} tailored to each district challenge.
\item \textbf{Expanded fail-forward examples} for tricky bridge steps.
\item \textbf{Position shift guidelines} for \Diamond{} usage.
\end{itemize}

\section{XP Progression \& Milestones (Total: 54--90 XP)}
\label{app:xp}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Milestone} & \textbf{XP Award} & \textbf{Triggers When}\\
\midrule
Bridge Completion & +12 XP & Entire party reaches Seven-Sorrows\\
District Solution & +6 XP & Each of first 3 districts resolved\\
Sorrow Mastery & +8 XP & 4+ districts solved or major echo pacified\\
Orrery Access & +10 XP & Chambers unlocked (before finale)\\
Final Resolution & +15 XP & Orrery outcome achieved\\
Session Base & +6 XP & Per session (exploration, roleplay, setbacks)\\
\bottomrule
\end{tabularx}

\section{Magic Path Integration by District}
\label{app:magic-by-district}
\subsection*{District-Specific Advantages}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{District} & \textbf{Magic Path Benefits}\\
\midrule
\textbf{Broken Promises} & \textit{Runekeepers:} Patron's authority can compel door compliance (+1 Effect). \textit{Invokers:} Symbols reveal which oaths were to which patrons.\\
\textbf{Final Regrets} & \textit{Casters:} Elemental magic can reshape reenactments (Fire for passion, Water for acceptance). \textit{Cantors:} Songs provide +1 die to alternative endings.\\
\textbf{Unfinished Songs} & \textit{Cantors:} Auto-success on melody completion if theme matches patron. \textit{Runekeepers:} Rites can harmonize with city's rhythm (reduce Rift Instability).\\
\textbf{Unsent Letters} & \textit{Invokers:} Symbols authenticate letters as ``patron-approved.'' \textit{Casters:} Air magic can deliver letters instantly (but attracts echoes).\\
\textbf{No Farewell} & \textit{Runekeepers:} Patron favor allows safe pruning (+1 Position). \textit{Summoners:} Spirits can test entanglement without gear loss.\\
\textbf{Abandoned Vows} & \textit{All Paths:} Vow-related magic gains +1 Effect. Breaking a magical vow here costs double Obligation/Backlash.\\
\textbf{Drowned Forum} & \textit{Casters:} Water magic provides breathing advantage. \textit{Runekeepers:} Storm/Sea patrons grant debate leverage.\\
\bottomrule
\end{tabularx}

\section{Fail-Forward Reference (Bridge Steps)}
\label{app:fail-forward}
\subsection*{Expanded Consequences}
\begin{tabularx}{\linewidth}{@{}l X X@{}}
\toprule
\textbf{Step} & \textbf{Partial Success (1 Boon)} & \textbf{Miss (2 Boons + Cost)}\\
\midrule
Promise Binding & Recall who you made it to; gain +1 die vs that person & Vow becomes audible to rivals; they can use it against you\\
Price of Knowledge & Learn the question's origin; gain lore advantage & Foresee a comrade's potential demise; Dread +2\\
Unburied Failure & Echo appears but offers useful memories & Echo merges with a rival; they gain your mannerisms\\
Overused Tool & Recognize dependency; next use gains +1 Effect & Tool becomes sentient and refuses certain tasks\\
Self-Unforgiveness & Apology to party member clears Dread immediately & Project failure onto ally; they suffer your consequence\\
Save One Thing & Item becomes indestructible for this journey & Rival now possesses it and knows its significance to you\\
\bottomrule
\end{tabularx}

\section{Diamond Position Shifts --- Quick Reference}
\label{app:diamonds}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Diamond} & \textbf{Position Shift} & \textbf{Limitations}\\
\midrule
Guide's Favor & Desperate $\rightarrow$ Controlled & Once per bridge segment; must help another traveler\\
Right of Passage & Risky $\rightarrow$ Dominant & Single portal; expires at sunset\\
Bearer's Seal & Controlled $\rightarrow$ Dominant (social) & Only for delivery missions; recipient must be living\\
Shared Vow & Group +1 Position step & All participants must uphold vow terms\\
Clear Conscience & Ignore taboo penalty & One use; must be narratively appropriate\\
Anchored Footing & Immune to \SB{} splits & Until next toll or camp\\
\bottomrule
\end{tabularx}

\section{Part I Aids --- The Bridge of a Thousand Steps (Tier II)}
\label{app:bridge-aids}
\subsection*{Enhanced Magic Interactions}
\textbf{Step Solutions by Path:}
\begin{itemize}
\item \textit{Runekeepers:} Invoke patron rites to answer steps---counts as ``deed'' but marks +1 Obligation.
\item \textit{Casters:} Elemental magic can create temporary handholds/footholds---provides +1 die but risks backlash on 1s.
\item \textit{Invokers:} Symbol display allows ritual answers---slower but safer (no Toll on failure).
\item \textit{Cantors:} Songs can answer melodic steps automatically---but attract sorrow-echo audience.
\item \textit{Summoners:} Spirits can test steps first---Leash ticks instead of PC paying cost.
\end{itemize}

\subsection*{XP Triggers --- Part I}
\begin{itemize}
\item \textbf{Each 2 steps solved}: +2 XP (max +10 XP for bridge)
\item \textbf{Philosopher's Truss success}: +4 XP
\item \textbf{Toll paid meaningfully}: +3 XP (confession/scar $>$ coin)
\item \textbf{Mirrored Step resolved}: +3 XP per pair
\item \textbf{Bridge completion}: +12 XP (total Part I: $\sim$32 XP)
\end{itemize}

\section{Part II Aids --- The Sunken City of Seven-Sorrows (Tier III--IV)}
\label{app:sorrows-aids}
\subsection*{Magic Path Finale Advantages}
\textbf{Orrery Pillars --- Path Specializations:}
\begin{itemize}
\item \textbf{Axis of Names:} Runekeepers/Invokers gain +2 dice (patron connections).
\item \textbf{Weights of Debt:} Casters can manipulate with elemental balance (+1 Effect).
\item \textbf{Gnomon of Return:} Cantors/Summoners align with rhythm (+1 Position).
\end{itemize}

\subsection*{XP Triggers --- Part II}
\begin{itemize}
\item \textbf{Each district solved}: +6 XP (max +42 XP for 7 districts)
\item \textbf{Rival setback}: +4 XP per major rival obstacle overcome
\item \textbf{Rift stabilization}: +8 XP (per successful surge calming)
\item \textbf{Orrery access}: +10 XP
\item \textbf{Finale resolution}: +15 XP (total Part II: $\sim$75 XP)
\end{itemize}

\section{Session Planning with XP --- Sample 3-Session Arc}
\label{app:sessions}
\subsection*{Session 1 (Bridge Start): $\sim$18 XP}
\begin{itemize}
\item 4 steps solved: +8 XP
\item Philosopher's Truss: +4 XP
\item Session base: +6 XP
\end{itemize}

\subsection*{Session 2 (Bridge End): $\sim$20 XP}
\begin{itemize}
\item Bridge completion: +12 XP
\item Meaningful toll: +3 XP
\item Session base: +6 XP
\end{itemize}

\subsection*{Session 3 (First District): $\sim$16 XP}
\begin{itemize}
\item District solved: +6 XP
\item Rival obstacle: +4 XP
\item Session base: +6 XP
\end{itemize}

\section{Enhanced NPC Magic Interactions}
\label{app:npc-magic}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{NPC} & \textbf{Magic-Specific Hooks}\\
\midrule
Captain Brann & Respects Runekeepers with warlock patrons; distrusts ``wild'' casters\\
Sister Ysol & Offers sacraments that function as temporary \Diamond{}s for religious magic users\\
Cartographer Pell & Seeks magical mapping assistance; pays in ritual components\\
Toll-Taker & Recognizes patron bonds; may waive tolls for respectful invokers\\
Conductor of Songs & Recruits cantors for city harmonies; offers musical lore\\
\bottomrule
\end{tabularx}

\section{Player-Facing Magic Guidelines}
\label{app:player-magic}
\textbf{For Casters:} Your magic reacts strongly to sorrow-echoes. Backlash may manifest as amplified regrets or temporal echoes.

\medskip
\textbf{For Runekeepers:} Your patron watches this place closely. Obligation marks may come with visions of the city's past patrons.

\medskip
\textbf{For Invokers:} Symbols hum with recognition here. Rituals take half time but attract echo-witnesses.

\medskip
\textbf{For Cantors:} Unfinished songs seek completion. Your performances may temporarily heal rift instability.

\medskip
\textbf{For Summoners:} Spirits behave oddly here. They may remember things from the city's past or refuse certain commands.

\vfill
\begin{center}
\small{\emph{Steps Into Sorrow} --- Appendix Edition}
\end{center}

\end{document}