\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage{enumitem}
\hypersetup{colorlinks=true, linkcolor=black, urlcolor=blue}

\newcommand{\Part}[1]{\section*{#1}}
\newcommand{\Clock}[2]{\textbf{#1} [#2]}
\newcommand{\Term}[1]{\textbf{#1}}
\newcommand{\DV}{\Term{DV}}
\newcommand{\Position}{\Term{Position}}
\newcommand{\SB}{\Term{SB}}
\newcommand{\Diamond}{\Term{Diamond}}
\newcommand{\Tier}{\Term{Tier}}

\definecolor{softgray}{RGB}{245,245,245}
\definecolor{inkblue}{RGB}{20,40,120}
\definecolor{blood}{RGB}{130,10,10}
\definecolor{forest}{RGB}{16,100,60}

\tcbset{colback=softgray, colframe=black, boxrule=0.4pt, arc=2mm}

\begin{document}

\begin{center}
{\LARGE \textbf{Stone \& Silence}}\\[4pt]
{\large A Two-Part Horror Arc for \emph{Fate's Edge}}\\[2pt]
\textit{Tiers II--IV \;|\; 5--8 sessions \;|\; Nonverbal rites, reality distortion, art-as-apocalypse}
\end{center}

\hrule
\vspace{0.5em}

\section*{Using This Packet}
This packet presents \emph{Stone \& Silence} primarily as a suite of appendices to slot alongside the core write-ups for:
\begin{itemize}
\item Part I: \textbf{The God That Feeds on Silence} (Tier II, monastery heist-of-ritual).
\item Part II: \textbf{The Star-Mad Sculptor of Weeping Crag} (Tier III--IV, art-driven catastrophe).
\end{itemize}
Each appendix is a ready reference for the table: milestones, magic options, fail-forward tables, Diamonds, and quick sheets.

\appendix

\section{GM Enhancements Overview}
\label{app:enhancements}
\textbf{Included options:}
\begin{itemize}
\item \textbf{XP milestone awards} tuned to a 5--8 session arc.
\item \textbf{Magic path specializations} for the silence-rite and star-geometry play spaces.
\item \textbf{Expanded fail-forward tables} for vow puzzles and art hazards.
\item \textbf{Position shift guidelines} for \Diamond{} usage and nonverbal leverage.
\end{itemize}

\section{XP Progression \& Milestones (Approx.\ 48--84 XP)}
\label{app:xp}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Milestone} & \textbf{XP Award} & \textbf{Triggers When}\\
\midrule
Peaceful Entry to Monastery & +6 XP & Gain access without breaking vow spaces\\
Rite of Discord Prepared & +8 XP & Components assembled; rhythm mapped\\
Sound Loss Stabilized & +10 XP & Sound returns to a wing without harm\\
Trail to Weeping Crag Secured & +6 XP & Passage, supplies, or local pact obtained\\
Decode Two Sculptures & +8 XP & Interpret geometry and its hazard\\
Confront Elara & +10 XP & Debate or duel of wills begins\\
Final Resolution & +15 XP & Entity banished, bargained, or contained\\
Session Base & +6 XP & Per session (investigation, play, setbacks)\\
\bottomrule
\end{tabularx}

\section{Magic Path Integration by Site}
\label{app:magic-by-site}
\subsection*{Monastery of the Mute (Silence God)}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Path} & \textbf{Benefits \& Risks}\\
\midrule
Cantors & Conductors of nonverbal cadence; can ``sing'' in sign. \emph{Benefit:} once/scene reduce \DV{} by 1 when leading mime or cadence. \emph{Risk:} echo-choirs mimic you, ticking \Clock{Sound Loss}{+1}.\\
Runekeepers & Seal-writing and vow-binding. \emph{Benefit:} sanctify boundaries to block the entity's feeding. \emph{Cost:} +1 Obligation per wing protected.\\
Invokers & Symbolic intercession with abstract entities. \emph{Benefit:} distract or redirect the Audient Void briefly. \emph{Risk:} symbols ``deafen'' an ally for a beat.\\
Casters & Elemental hush or discord. \emph{Benefit:} create safe noise pockets for rites. \emph{Risk:} backlash manifests as silence zones following you.\\
Summoners & Silent servitors and sign-bound spirits. \emph{Benefit:} test taboos by proxy. \emph{Risk:} spirit interprets vows literally at a bad moment.\\
\bottomrule
\end{tabularx}

\subsection*{Weeping Crag Atelier (Star-Shaper)}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Path} & \textbf{Benefits \& Risks}\\
\midrule
Cantors & Counter-harmonies to spatial chords. \emph{Benefit:} lower one Reality DV by 1. \emph{Risk:} dissonance attracts shards.\\
Runekeepers & Null sigils over geometry. \emph{Benefit:} freeze a distortion for 1 beat. \emph{Cost:} sigil burns to ash when touched by starlight.\\
Invokers & Parley with concept-aspect. \emph{Benefit:} bargain an angle closed. \emph{Risk:} a new, lesser angle opens elsewhere.\\
Casters & Geometric evocations. \emph{Benefit:} anchor a safe Euclidean bubble. \emph{Risk:} bubble drifts, splitting the party.\\
Summoners & Sculptor's muses, bound. \emph{Benefit:} extract meaning from a piece. \emph{Risk:} muse compels creation during rest.\\
\bottomrule
\end{tabularx}

\section{Fail-Forward Reference}
\label{app:fail-forward}
\subsection*{Monastery --- Silence Rites \& Taboos}
\begin{tabularx}{\linewidth}{@{}l X X@{}}
\toprule
\textbf{Test} & \textbf{Partial Success (1 Boon)} & \textbf{Miss (2 Boons + Cost)}\\
\midrule
Gesture-Only Debate & You sway one elder; lower next social \DV{} by 1 & A forbidden word slips; \Clock{Sound Loss}{+1} and vow penance\\
Forbidden Words Maze & Map 70\% of safe routes & You mark a safe route but swap two glyphs; escort is misled\\
Rite of Discord & You craft an off-beat version; workable & Rite works but stuns a wing; allies Deafened (1 scene)\\
Silent Vigil & Hold most of the watch & Entity tastes your breath; shadow follows until named\\
Library Search & Find two of three volumes & Missing folio lies with a novice who now idolizes you\\
\bottomrule
\end{tabularx}

\subsection*{Weeping Crag --- Art Hazards \& Angles}
\begin{tabularx}{\linewidth}{@{}l X X@{}}
\toprule
\textbf{Test} & \textbf{Partial Success (1 Boon)} & \textbf{Miss (2 Boons + Cost)}\\
\midrule
Reality Distortion Field & Stabilize a corridor for a beat & Corridor inverts gravity later; party splits\\
Reading the Piece & Extract theme; lower \DV{} by 1 & Take the idea into you; Disadvantage on rest until purged\\
Studio Debate with Elara & Earn a hesitation & Elara adopts your logic; \Clock{Elara's Conviction}{+1}\\
Shattering a Seed & Break a growth node & Shards seed a rival camp; \Clock{Town Panic}{+1}\\
Euclidean Bubble & Anchor a sphere & Bubble drifts to the cliff edge; equipment test or loss\\
\bottomrule
\end{tabularx}

\section{Diamond Position Shifts --- Quick Reference}
\label{app:diamonds}
\begin{tabularx}{\linewidth}{@{}l l X@{}}
\toprule
\textbf{Diamond} & \textbf{Position Shift} & \textbf{Limitations}\\
\midrule
Silent Vow & Desperate $\rightarrow$ Controlled in vow spaces & Broken by voluntary speech; silent sign allowed\\
Abbot's Seal & Risky $\rightarrow$ Dominant when invoking law & Expires at sundown; one wing only\\
Bell of Discord & Lower ritual \DV{} by 1 when rung off-key & Each ring ticks \Clock{Sound Loss}{+1}\\
Artist's Proof & Dominant versus a specific sculpture & Proof burns on use; one piece only\\
Resonance Diagram & Group +1 \Position{} against distortions & Consumed if you flee\\
Euclidean Anchor & Immune to one geometry SB split & Until the anchor warms (1 scene)\\
\bottomrule
\end{tabularx}

\section{Part I Aids --- The God That Feeds on Silence (Tier II)}
\label{app:part1}
\subsection*{Core Clocks}
\begin{itemize}
\item \Clock{Sound Loss}{6} (areas fall permanently silent)
\item \Clock{Library Taboos}{4} (forbidden words pressure)
\item \Clock{Monastic Trust}{4} (your standing with elders)
\end{itemize}

\subsection*{Scene Menu}
\paragraph{Gate of Palms.} Gesture-only parley for entry. \Position{} starts Controlled; Cantor-led sign can drop \DV{} by 1.
\paragraph{Forbidden Words Stacks.} Navigate aisles where speaking the wrong lexeme seals shelves; map with chalk and mime.
\paragraph{Rite of Discordant Noise.} Assemble a counter-cadence to feed and redirect the entity; success shifts \Clock{Sound Loss}{-2}.
\paragraph{The Silent Bell.} A bell that rings as absence; ring off-key to ``feed'' the entity away while risking a vow breach.

\subsection*{\SB{} Spend Prompts}
\begin{itemize}
\item Echo-silence swallows a sentence; repeat the action or pay time.
\item Novice imitates you; a taboo is broken in your name.
\item A wing loses ambient sound; stealth gains, social loses.
\end{itemize}

\subsection*{XP Triggers --- Part I}
\begin{itemize}
\item \textbf{Peaceful Entry}: +6 XP
\item \textbf{Map 3 Safe Routes}: +4 XP
\item \textbf{Stabilize a Wing}: +6 XP
\item \textbf{Complete the Rite}: +8 XP
\item \textbf{Session Base}: +6 XP
\end{itemize}

\section{Part II Aids --- The Star-Mad Sculptor of Weeping Crag (Tier III--IV)}
\label{app:part2}
\subsection*{Core Clocks}
\begin{itemize}
\item \Clock{Local Reality Collapse}{8}
\item \Clock{Star Geometry Spread}{6}
\item \Clock{Town Panic}{4}
\item \Clock{Elara's Conviction}{6}
\end{itemize}

\subsection*{Set-Pieces}
\paragraph{The Cantilevered Walk.} A cliff path where angles disagree. Success allows a shortcut; cost drops supplies into a sideways crevasse.
\paragraph{The Studio Debate.} Persuade Elara that beauty must spare the world. Break her tie to the entity by changing her aesthetic premise.
\paragraph{Shattering the Seed.} Destroy a growth node before it harmonizes; risk shard swarms and rival interference.
\paragraph{Night of Fixed Stars.} The sky locks; time dilates; use Diamonds to keep \Position{} stable while dismantling pillars.

\subsection*{Ritual Pillars (Break Any One)}
\begin{itemize}
\item \textbf{Keening Axis:} The tone line that holds the transformation. \emph{Counter:} ring the Bell of Discord off cadence.
\item \textbf{Canvas Lattice:} The pattern net in the atelier. \emph{Counter:} overlay a Resonance Diagram to introduce noise.
\item \textbf{Nine Angles:} The geometric premises in Elara's manifesto. \emph{Counter:} refute an angle in public; shift \Clock{Elara's Conviction}{-2}.
\end{itemize}

\subsection*{\SB{} Spend Prompts}
\begin{itemize}
\item Gravity flips on a beat; split the party or drop equipment.
\item Shards mimic a PC's gesture; their action repeats at the worst time.
\item A rival ``secures'' a sculpture you needed; pay a price to study it.
\end{itemize}

\subsection*{XP Triggers --- Part II}
\begin{itemize}
\item \textbf{Decode a Sculpture}: +4 XP (max 3 times)
\item \textbf{Stabilize a Distortion}: +6 XP
\item \textbf{Persuade or Stop Elara}: +10 XP
\item \textbf{Final Resolution}: +15 XP
\item \textbf{Session Base}: +6 XP
\end{itemize}

\section{Session Planning with XP --- Sample 3-Session Arc}
\label{app:sessions}
\subsection*{Session 1 (Monastery): $\sim$18 XP}
\begin{itemize}
\item Peaceful Entry: +6 XP
\item Map 3 Safe Routes: +4 XP
\item Session Base: +6 XP
\end{itemize}

\subsection*{Session 2 (Rite \& Trail): $\sim$16 XP}
\begin{itemize}
\item Stabilize a Wing: +6 XP
\item Complete the Rite: +8 XP
\item Session Base: +6 XP
\end{itemize}

\subsection*{Session 3 (Weeping Crag): $\sim$20--24 XP}
\begin{itemize}
\item Decode a Sculpture: +4 XP
\item Stabilize a Distortion: +6 XP
\item Persuade/Stop Elara: +10 XP
\item Session Base: +6 XP
\end{itemize}

\section{Enhanced NPC Magic Interactions}
\label{app:npc}
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{NPC} & \textbf{Magic-Specific Hooks}\\
\midrule
Abbess Mirel & Trades silence hours as Diamonds; favors Runekeepers who seal taboos.\\
Brother Slate & Cantor of the Hands; teaches sign-cant that lowers \DV{} once/scene.\\
Elara of Weeping Crag & Will debate any premise; grants Artist's Proof to those who impress her.\\
Foreman Jote & Quarry boss; knows how the angles ``bite''; lends Euclidean Anchors for a price.\\
Rival Adept Sera & Invoker studying the Star-Shaper; might help you close an angle, for a debt.\\
\bottomrule
\end{tabularx}

\section{Player-Facing Guidelines}
\label{app:player}
\textbf{For All:} Position is everything. Earn Dominant by securing Diamonds first; roll later.

\medskip
\textbf{For Casters:} Backlash manifests as lost senses or skewed geometry; treat it as fiction, not a simple penalty.

\medskip
\textbf{For Runekeepers:} Your seals can anchor safety, but the entity hunts boundaries; expect targeted pressure.

\medskip
\textbf{For Invokers:} Parley with abstract aspects is possible; be ready to give up a symbol or habit to make it stick.

\medskip
\textbf{For Cantors:} Nonverbal music is still music. Conduct in silence; let hands be the choir.

\medskip
\textbf{For Summoners:} Spirits will ``remember'' wrong; let them test hazards, but mind their literalism.

\section{Printable Clocks}
\label{app:clocks}
\begin{tcolorbox}
\Clock{Sound Loss}{6}\hfill\(\square\square\square\square\square\square\)\\[3pt]
\Clock{Library Taboos}{4}\hfill\(\square\square\square\square\)\quad
\Clock{Monastic Trust}{4}\hfill\(\square\square\square\square\)\\[3pt]
\Clock{Local Reality Collapse}{8}\hfill\(\square\square\square\square\square\square\square\square\)\\[3pt]
\Clock{Star Geometry Spread}{6}\hfill\(\square\square\square\square\square\square\)\quad
\Clock{Town Panic}{4}\hfill\(\square\square\square\square\)\quad
\Clock{Elara's Conviction}{6}\hfill\(\square\square\square\square\square\square\)
\end{tcolorbox}

\vfill
\begin{center}
\small{\emph{Stone \& Silence} --- Appendix Edition}
\end{center}

\end{document}