\documentclass[12pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{mdframed}
\usepackage{needspace}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{calc}
\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{longtable}
\usepackage{pdflscape}
\usepackage{multicol}
\usepackage{graphicx}

\definecolor{darkblue}{rgb}{0.0, 0.0, 0.55}
\hypersetup{
    colorlinks=true,
    linkcolor=darkblue,
    urlcolor=darkblue,
    pdftitle={The Tithe of Forgotten Faces},
    pdfauthor={Fate's Edge Adventure Module}
}

% ------------------- Custom commands -------------------
\newcommand{\tag}[1]{\textsc{#1}}
\newcommand{\clk}[1]{\textit{[#1]}}
\newcommand{\stringitem}[1]{\textit{#1}}
\newcommand{\dusana}[1]{%
    \begin{quote}\itshape #1 ---~Dusana of the Raven Road\end{quote}}
\newcommand{\letter}[2]{%
    \begin{quote}\small\itshape #1 \hfill --- #2\end{quote}}

% ------------------- Styling -------------------
\titleformat{\chapter}{\Huge\bfseries}{\thechapter}{20pt}{}
\titleformat{\section}{\Large\bfseries}{\thesection}{15pt}{}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8ex}

% ------------------- Infobox styling -------------------
\newmdenv[
    linecolor=black,
    backgroundcolor=gray!5,
    innertopmargin=8pt,
    innerbottommargin=8pt,
    innerrightmargin=8pt,
    innerleftmargin=8pt,
    skipabove=6pt,
    skipbelow=6pt,
    font=\small
]{infobox}

% ------------------- Document -------------------
\begin{document}

\title{
    \Huge\textbf{The Tithe of Forgotten Faces}\\[0.5em]
    \Large A Fate's Edge Adventure
}
\author{
    Tier II--III (2--3 Sessions)\\
    \small Viterra Hedge-Law Series
}
\date{
    \small Written in the margin of the Ledger
}
\maketitle

\tableofcontents

\chapter{Overview: The Law of Thornwood}

The law of Viterra is written in hedge and precedent. But some debts are not recorded in any ledger. They are etched in the spaces where a face used to be.

\vspace{1em}
\dusana{I have seen courts judge theft of grain, theft of land, even theft of breath. But theft of memory? That is a crime without a body. The victim stands before you, breathing, blinking, yet missing the very thing that makes them a person. Who owns the past? The one who lived it, or the one who remembers it?}

\section{Premise}

A woman stands accused before a Viterran hedge-court of \textbf{hollowing} her husband---erasing his memory of their marriage using a forbidden rite. The party serves as jurors, investigators, and ultimately judges. But the husband, now a blank-eyed stranger, claims he \textit{asked} to be forgotten. The court's ledger is a magical artifact that can restore or excise memories at a price. The party must uncover the truth: was this a crime, an act of mercy, or something else entirely?

\section{Themes}

\begin{itemize}
    \item \textbf{Memory as Property:} Can a memory be owned, traded, or stolen?
    \item \textbf{The Right to Be Forgotten:} Is oblivion a mercy or a violation?
    \item \textbf{The Violence of Care:} When does protection become control?
    \item \textbf{The Weight of Witness:} What obligation do you have to the truth you uncover?
\end{itemize}

\section{Tone}

Quiet horror, legal drama, moral ambiguity. The horror is not in gore, but in absence. The drama is not in shouting, but in silence. The ambiguity is not in the evidence, but in the interpretation of consent.

\section{Key Mechanics}

\begin{infobox}
\textbf{Adventure Mechanics}
\begin{itemize}
    \item \textbf{Memory as a Resource:} Personal memories can be offered as leverage, lost as costs, or stolen by entities.
    \item \textbf{Ledger of Recollection:} A magical artifact that can restore or remove memories, but each use accrues a Debt.
    \item \textbf{Verdict Clock:} Tracks the court's patience and the case's complexity.
    \item \textbf{Personalization:} Each PC is asked early on: \textit{What memory would you least want to lose?} That memory becomes a mechanical token that can be spent or stolen.
\end{itemize}
\end{infobox}

\section{Locations}

\begin{itemize}
    \item \textbf{Thornwood Hedge Court:} A ring of standing stones draped in moss, with a central slate table.
    \item \textbf{The Accused's Cottage:} Smells of dried thyme and old paper. Cluttered with herbs.
    \item \textbf{The Hospice of St. Jude:} Where Corin currently resides. Quiet, sterile, filled with humming.
    \item \textbf{The Well of Echoes:} A hidden spring where memories can be drowned. Water is black and still.
\end{itemize}

\section{Key NPCs}

\begin{itemize}
    \item \textbf{Magistra Mira Venn:} The Queen's Justicar. Stern but fair. Wears gray robes and carries a brass key.
    \item \textbf{Anya of Thornwood:} The accused. A hedge-witch in her fifties. Weary, eyes carrying the weight of a choice.
    \item \textbf{Corin the Hollow:} The husband. Empty of memory, speaks in third person, eerily polite.
    \item \textbf{Brother Tallow:} A Witness cultist. Wants to use the case to outlaw memory magic altogether.
\end{itemize}

\section{The Hook}

The party is summoned by Magistra Venn as \textbf{neutral witnesses} known for their discretion. The case has already drawn the attention of the Witness cult, who see it as a chance to set precedent. The party arrives at the crossroads outside Thornwood at dawn.

\chapter{Act I: The Hollow Man}

\section{Scene 1: The Summons}

The morning mist clings to the hedgerows like old linen. Magistra Mira Venn meets the party at the crossroads, a ledger under one arm and a bell in her hand.

\letter{Anya of Thornwood stands accused of unmaking her husband's mind. The court convenes at noon. You are not to judge---yet. You are to witness and advise. The law of Viterra has no precedent for this.}{--- Magistra Mira Venn}

The party is given a writ of passage (a Diamond: \textit{Witness Seal}). They are told they may examine the accused, the hollowed husband, and any evidence before the court sits.

\textbf{Keeper Instructions:} Establish the stakes immediately. Ask each player: \textit{What is one memory your character would least want to lose?} Write this down on a card. This becomes their \textbf{Memory Token}.

\section{Scene 2: The Accused}

Anya's cottage is warm but dim. She is a hedge-witch in her fifties, hands stained with herbs. She does not deny that she removed Corin's memories.

\textbf{Dialogue:} \textit{He was drowning in them. Every day he woke and relived our worst argument, the child we lost, the war he fled. I did not hollow him. I pruned him. He asked me to, when he still had a voice.}

She offers the party her own memory of that conversation as a \textbf{Memory Token}. If they accept, they experience it firsthand (a flashback scene). The memory is raw, painful, and ambiguous: Corin did say \textit{Make me forget}, but was it consent or despair?

\textbf{Mechanic:} Accepting Anya's Memory Token grants +1 Die to all \textbf{Insight} rolls regarding the case, but marks 1 \textbf{Fatigue} (the emotional weight).

\section{Scene 3: The Hollow}

Corin sits in a rocking chair at the Hospice of St. Jude, humming a tune he does not recognize. He is gentle, vacant, and polite.

\textbf{Dialogue:} \textit{She smells of thyme. I think I loved her. But I don't know why.}

He does not accuse her. He does not defend her. He is a living absence.

\textbf{Complication:} Brother Tallow is already there, taking notes. He argues: \textit{Memory is the foundation of personhood. To erase it is murder by inches. The law must name this as soul-theft.} He offers the party a bribe (a Diamond: \textit{Cult Favor}) to testify against Anya. Accepting the bribe grants leverage but marks 1 \textbf{Obligation} to the Witness Cult.

\section{Scene 4: The Ledger of Recollection}

Magistra Venn shows the party the court's artifact: a brass-bound book that writes memories as they are spoken. It can also extract them, leaving the speaker momentarily blank.

\begin{infobox}
\textbf{The Ledger's Function}
\begin{itemize}
    \item \textbf{Restore:} Can restore Corin's memory.
    \item \textbf{Cost:} Someone must offer a memory of equal weight to the one restored.
    \item \textbf{Constraint:} The court cannot compel that sacrifice---it must be voluntary.
    \item \textbf{Debt:} Using the Ledger incurs a \textit{Ledger Debt}. The user forgets something random (GM picks a minor fact, like a password or a face).
\end{itemize}
\end{infobox}

This sets up the final moral choice. The party must determine if restoration is possible, desired, or just.

\chapter{Act II: The Trial of What Remains}

The court sits at dusk. The standing stones hum with old law. The party is given the role of \textit{amici curiae}---friends of the court. They may question witnesses, examine evidence, and propose verdicts.

\section{Key Witnesses}

\begin{itemize}
    \item \textbf{The Neighbor (Old Marta):} Saw Anya performing a rite in the garden with a silver bowl and a hair from Corin's head. She believes Anya is a witch in the old, dangerous sense.
    \item \textbf{The Apothecary (Guild Master Venn):} Testifies that Corin came to him a year ago, asking for \textit{something to quiet the nightmares}. He refused, as it was beyond his craft.
    \item \textbf{The Husband's Brother (Farin):} Demands punishment. \textit{He is a shell. She should be exiled.}
    \item \textbf{Anya's Apprentice (Lena):} Secretly gives the party a letter Corin wrote before the hollowing. It says: \textit{If I lose myself, do not blame her. Blame the war. Blame the child who did not live. But let me go.}
\end{itemize}

\section{Timers}

\begin{infobox}
\textbf{Trial Timers}
\begin{itemize}
    \item \textbf{Court's Patience [4]:} Advances each time the party introduces irrelevant evidence or fails to move the case forward. When full, Magistra Venn issues a default ruling (guilty, with standard punishment of exile).
    \item \textbf{Memory Fracture [6]:} Tracks the instability of the Ledger. Advances each time it is used to extract or restore a memory.
    \begin{itemize}
        \item \textbf{At 3 segments:} The Ledger begins to leak---random memories from past cases appear as ghostly echoes.
        \item \textbf{At 6 segments:} It must be sealed for a year, losing all unrecorded cases.
    \end{itemize}
\end{itemize}
\end{infobox}

\section{Mechanics of Testimony}

PCs can use \textbf{Presence + Sway} to coax testimony, \textbf{Wits + Insight} to detect lies, or \textbf{Spirit + Lore} to interpret magical evidence.

\begin{itemize}
    \item \textbf{Success:} Grants a \textbf{Truth Token} (a Boon that can be spent to influence the final verdict).
    \item \textbf{Failure:} Advances \textbf{Court's Patience} by 1.
    \item \textbf{Critical Failure (1):} Advances \textbf{Court's Patience} by 2 and reveals a confusing contradiction.
\end{itemize}

\chapter{Act III: The Well of Echoes}

Midnight. The court adjourns, but Magistra Venn asks the party to investigate a hidden location: the Well of Echoes, where Anya says she \textit{drowned} Corin's memories. It is a spring deep in the woods, ringed with white stones. The water is black and still, and when you look into it, you see not your reflection but fragments of forgotten moments---not yours, but others'.

\section{The Encounter: Memory Eater}

The Well is guarded by a \textbf{Memory Eater} (a spirit summoned by past use of the Ledger). It attacks not with harm but with theft.

\begin{infobox}
\textbf{Memory Eater Stats}
\begin{itemize}
    \item \textbf{Armor:} None (Immune to physical harm).
    \item \textbf{Attack:} \textit{Mind Drain}. Each successful attack erases one of the PC's Memory Tokens (or, if none remain, a random memory of the GM's choice).
    \item \textbf{Defense:} If the party tries to fight it, it multiplies---each blow creates another copy.
    \item \textbf{Weakness:} The only way to pass is to \textbf{offer} a memory willingly---something precious, given to the Well as payment for access.
\end{itemize}
\end{infobox}

This is where the game gets personal. The GM should have asked each PC in Act I: \textit{What memory would you least want to lose?} Now they may have to give it up to learn the truth.

\section{The Truth}

At the bottom of the Well, the party finds not memories, but a single object: a locket with a faded portrait of a child. Anya and Corin's daughter, who died of fever. The grief was the root of Corin's unraveling.

\textbf{Revelation:} Anya did not erase that grief---she \textit{took it into herself}, using the Ledger's power at great personal cost. The evidence is in her own memory: she has not slept peacefully in three years; she sees the child's face every night. She hollowed herself to save him.

\chapter{Act IV: The Verdict}

The court reconvenes at dawn. The party must advise Magistra Venn on the verdict.

\section{Verdict Options}

\begin{enumerate}
    \item \textbf{Guilty of Soul-Theft:} Exile Anya, destroy the Ledger. Sets a harsh precedent. Brother Tallow and the Witness cult are pleased; memory magic becomes a capital crime.
    \item \textbf{Not Guilty by Mercy:} Rule that consent, given in extremis, is still consent. Anya is free, but the Ledger is confiscated by the Crown. The cult is furious; a seed of rebellion is planted.
    \item \textbf{A Third Path (Restoration):} Use the Ledger to restore Corin's memories, but the cost is that someone (a PC or Anya) must sacrifice a memory of equal weight. The party must choose who pays. If a PC volunteers, they lose the memory they named in Act I (or another if they have none). This restores Corin but leaves a scar on the volunteer.
    \item \textbf{Rewrite the Law:} Argue that Viterran law has no provision for this case, and therefore a new precedent must be set: \textit{the right to be forgotten, if witnessed and voluntary, is not a crime.} This requires a successful social roll (DV 5, Presence + Law) and spends all Truth Tokens accumulated. If successful, the case becomes a landmark ruling; Anya is pardoned, and the Ledger is kept under neutral custody.
\end{enumerate}

\section{Epilogue}

Based on the verdict, the tone shifts. Anya may be freed, exiled, or executed. Corin may remain hollow, be restored, or---if the restoration fails---become a permanent resident of the hospice, humming a tune that no one can name. The party's reputation echoes through Viterra: defenders of memory, or butchers of the mind.

\begin{infobox}
\textbf{Consequences}
\begin{itemize}
    \item \textbf{Exile:} Anya disappears into the wilds. The Ledger is broken. Memory magic goes underground.
    \item \textbf{Mercy:} Anya remains in Thornwood. The Ledger is secured. The Cult begins plotting against the Court.
    \item \textbf{Restoration:} Corin remembers the pain. He and Anya must rebuild their marriage on new ground. The PC who sacrificed remembers the cost.
    \item \textbf{New Law:} A legal precedent is set. Future cases of memory alteration are judged by consent, not outcome.
\end{itemize}
\end{infobox}

\chapter{Mechanics and Handouts}

\section{Memory Tokens}

Each PC receives a blank token card at the start. They write down one memory: a specific event, person, or sensation they would not want to lose. Examples: \textit{The smell of my mother's bread.} \textit{The face of my first love.} \textit{The day I swore my oath.}

\begin{infobox}
\textbf{Token Rules}
\begin{itemize}
    \item \textbf{Spent:} During the trial, spend to gain +1 die on a social or Lore roll---the memory is momentarily recalled with perfect clarity, then fades slightly.
    \item \textbf{Stolen:} By the Memory Eater or by the Ledger's misuse. Lost forever, unless restored by a similar sacrifice.
    \item \textbf{Offered:} Willingly to the Well or the Ledger to gain crucial information or save another.
    \item \textbf{Depletion:} If a PC has no tokens left, any further memory loss is permanent and narrative (GM's choice).
\end{itemize}
\end{infobox}

\section{The Ledger's Debt}

Using the Ledger of Recollection to extract or restore a memory always incurs a \textbf{Debt}: the user forgets something random (GM picks a minor fact, like a password or a face). The Ledger does not give without taking.

\section{Personal Whispers}

Prepare 2--3 personalized whispers for each PC, based on their background, to use in the Well scene or during tense testimony.

\begin{itemize}
    \item \textit{You remember a door you never opened. Behind it, your father is still alive.}
    \item \textit{The water shows you holding a knife. Whose blood is that?}
    \item \textit{You hear your own voice speaking a language you do not know.}
\end{itemize}

\chapter{Appendix: NPCs and Locations}

\section{NPC Profiles}

\begin{longtable}{p{0.25\textwidth} p{0.7\textwidth}}
\toprule
\textbf{Magistra Mira Venn} & \textbf{Role:} Judge, Arbiter \\
& \textbf{Stats:} Presence 4, Wits 3, Spirit 3 \\
& \textbf{Goal:} Uphold the law while preventing chaos. \\
& \textbf{Secret:} She once used the Ledger to forget her own daughter's death. \\
\midrule
\textbf{Anya of Thornwood} & \textbf{Role:} Accused, Hedge-Witch \\
& \textbf{Stats:} Spirit 4, Wits 3, Presence 2 \\
& \textbf{Goal:} Protect Corin from pain, even at cost of self. \\
& \textbf{Secret:} She carries the grief of the dead child in her own mind. \\
\midrule
\textbf{Corin the Hollow} & \textbf{Role:} Victim, Witness \\
& \textbf{Stats:} Body 2, Spirit 1 (Damaged), Presence 2 \\
& \textbf{Goal:} Find peace. \\
& \textbf{Secret:} Fragments of memory return in dreams. \\
\midrule
\textbf{Brother Tallow} & \textbf{Role:} Antagonist, Cultist \\
& \textbf{Stats:} Presence 3, Spirit 3, Wits 2 \\
& \textbf{Goal:} Outlaw all memory magic. \\
& \textbf{Secret:} He fears what he might have forgotten himself. \\
\bottomrule
\end{longtable}

\section{Location Details}

\textbf{Thornwood Hedge Court:}
Ancient standing stones arranged in a circle. Moss grows thick on the north side. The central slate table is cold to the touch.
\textit{Mechanic:} Social rolls within the circle gain +1 Position if truth is spoken.

\textbf{The Accused's Cottage:}
Small, stone-built. Smells of dried thyme and old paper. Shelves lined with jars of preserved herbs.
\textit{Mechanic:} Searching the cottage requires Wits + Notice (DV 3). Success finds Corin's letter.

\textbf{The Hospice of St. Jude:}
Quiet, sterile, filled with humming. Patients sit in rocking chairs.
\textit{Mechanic:} Spirit rolls here are at Disadvantage due to the psychic dampening.

\textbf{The Well of Echoes:}
A spring deep in the woods, ringed with white stones. Water is black and still.
\textit{Mechanic:} Looking into the water requires a Spirit test (DV 4) or gain 1 Fatigue from psychic feedback.

\chapter{Keeper Instructions}

\section{Ethics of Memory}

This adventure is about the ethics of memory. Do not present a clear \textit{right} answer. Let the players argue. The goal is not to solve the puzzle, but to weigh the cost.

\section{Emotional Stakes}

The Memory Tokens are the heart of the emotional stakes. When a PC offers their precious memory, let the table feel the weight. Do not trivialize it. Describe the loss vividly. \textit{You reach for the memory of your mother's face, but there is only a smooth stone where it used to be.}

\section{Visuals}

Use the Ledger's visuals: glowing ink, pages that turn themselves, echoes of voices that have been recorded. The Well should feel cold and deep. The Court should feel ancient and judgmental.

\section{The Memory Eater}

The Memory Eater is not a monster to be slain. If the party tries to fight it, it multiplies. The only way to pass is to give, not take. Encourage negotiation over combat.

\section{The Verdict}

The final verdict can be decided by a vote, not a roll. The social roll for \textit{Rewrite the Law} is only if the party wants to force a new precedent against the magistrate's leanings. If they argue well, they may not need to roll. Trust the table's decision.

\chapter{Colophon}

The hedge-court of Thornwood still stands. The ledger is sealed in a vault beneath the moot-stones. Anya lives in a cottage with a garden of thyme and rue. Corin---well, Corin chose. He asked to remember only one thing: the lullaby. Not the words. Just the tune.

\vspace{1em}
\dusana{I wrote their names in my own ledger, under \textit{Harmed} and \textit{Helped}. The balance is not settled. It never is. The ninth cup is for the ones who forgot. The ninth cup is also for the ones who chose to remember.}

\vspace{1em}
\letter{At a crossroads inn where no one asked her name.}{--- Dusana of the Raven Road}

\end{document}
