\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
\use0{hyperref}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{lipsum}

% Set page geometry
\geometry{a4paper, margin=1in}

% Define colors
\definecolor{headerblue}{RGB}{40, 50, 120}
\definecolor{sectionblue}{RGB}{60, 80, 180}
\definecolor{accentorange}{RGB}{210, 105, 30}
\definecolor{backgroundgray}{RGB}{240, 240, 240}

% Format section titles
\titleformat{\section}
  {\normalfont\Large\bfseries\color{sectionblue}}{\thesection}{1em}{}
\titleformat{\subsection}
  {\normalfont\large\bfseries\color{sectionblue}}{\thesubsection}{1em}{}

% Custom commands
\newcommand{\adventuretitle}[1]{\begin{center}\colorbox{headerblue}{\parbox{\textwidth}{\centering\color{white}\Huge\textbf{#1}}}\end{center}}
\newcommand{\keypoint}[1]{\par\noindent\textbf{\textcolor{accentorange}{#1}}}
\newcommand{\npctitle}[1]{\par\noindent\textbf{#1}\par}

% Remove page numbers
\pagestyle{empty}

\begin{document}

\adventuretitle{The Truth That Cannot Be Told}

\begin{center}
\textit{A Silent Choir Cult Adventure for Fate's Edge}
\end{center}

\section{Adventure Overview}

\keypoint{Premise:} A Cantor named Elara, once a respected scholar of Thepyrgos, has become the Chorus Leader of a Silent Choir cult that has taken root in the town of Veridian Crossroads. What began as a small circle of truth-seekers has grown into a community that enforces absolute honesty, with devastating consequences for those who cannot bear the truth.

\keypoint{Tone:} A mix of mystery, horror, and moral dilemma - the cult isn't inherently evil, but their pursuit of truth has destructive consequences.

\keypoint{Length:} 2-3 sessions

\keypoint{Themes:} Truth vs. Comfort, The Cost of Knowledge, Community vs. Isolation

\section{The Town of Veridian Crossroads}

Veridian Crossroads is a small farming town known for its annual Harvest Festival and the mysterious Veridian Stones that mark the town's boundaries. It's a place where secrets have been kept for generations - about the stones, about the land, and about the people who live there.

\subsection{Recent Changes}
\begin{itemize}[leftmargin=*]
    \item The Harvest Festival was canceled for the first time in 150 years
    \item Townsfolk whisper about "the truth-tellers" who now govern the town
    \item Visitors are turned away at the crossroads
    \item The town's market has transformed into a "Truth Circle" where all business is conducted with perfect honesty
\end{itemize}

\section{The Silent Choir Cult}

Led by Elara, the cult has grown to include 42 members - roughly a third of the town. They've restructured the town around their core beliefs:

\subsection{The Hierarchy}
\begin{itemize}[leftmargin=*]
    \item \textbf{Chorus Leader (Elara):} The corrupted Cantor who now leads the cult
    \item \textbf{First Truth-Tellers (12 members):} Tier 2-3 corruption - serve as chroniclers and enforcers
    \item \textbf{Seekers (28 members):} Tier 1 corruption - the rank-and-file members who detect deception
    \item \textbf{Silent Ones (2 members):} Those who left the cult but couldn't bear the truth, now living in exile
\end{itemize}

\subsection{Cult Practices}
\begin{itemize}[leftmargin=*]
    \item Daily "Truth Circles" where all town business is conducted with perfect honesty
    \item A ritual where new members drink from the "Chalice of Clarity" (a mixture of local herbs that heightens perception)
    \item A system of "Truth Debt" where lies must be repaid with honest acts
    \item Daily "Cleansing" where memories of deception are written down and burned
\end{itemize}

\section{Adventure Hook}

\keypoint{The Disappearance:} A merchant named Kaelen has gone missing from Veridian Crossroads. The PCs are hired by Kaelen's family to find him, with the promise of 500 gold for safe return.

\keypoint{Why It Matters:} Kaelen was on the verge of exposing the cult's activities to the authorities. His disappearance is the first sign that the cult has turned violent rather than simply enforcing truth.

\section{Key Locations \& Encounters}

\subsection{1. The Crossroads (Entry Point)}
The PCs are stopped at the town's entrance by two Seekers (Tier 1 corruption).

\keypoint{Mechanics:} They're given a "Truth Coin" - a copper token that must be shown to speak freely. Deception is impossible.

\keypoint{Encounter:} The PCs must navigate the Truth Circle to gather information. Deception is impossible, but strategic truth-telling is possible ("I'm here to help" vs. "I'm here to stop you")

\subsection{2. The Market Square (Truth Circle)}
The former market is now a circle of white stones where all business is conducted.

\keypoint{Key Detail:} Kaelen was last seen here arguing with the Chorus Leader.

\keypoint{Encounter:} The PCs must navigate the Truth Circle to gather information about Kaelen's disappearance.

\subsection{3. The Veridian Stones}
Ancient monoliths that surround the town.

\keypoint{Secret:} The stones are ancient truth-amplifiers. The cult's goal is to use them to force the entire region to speak only the truth.

\keypoint{Encounter:} The PCs discover hidden inscriptions that reveal why the cult has taken root here.

\subsection{4. The Chorus House (Cult Headquarters)}
A large home converted into the cult's meeting place.

\keypoint{Key Detail:} Kaelen is being held in the basement.

\keypoint{Encounter:} Confronting Elara and the cult leadership. The PCs must decide how to free Kaelen without triggering violence.

\section{The Cult's True Purpose}

The Silent Choir cult has discovered that the Veridian Stones are ancient truth-amplifiers. The cult's goal is to use the stones to force the entire region to speak only the truth, believing this will end all conflict.

\keypoint{The Dilemma:} The stones are working too well - they're amplifying not just spoken truth, but the truth of thoughts and memories, causing mental breakdowns in those with deep secrets.

\keypoint{Kaelen's Discovery:} He was going to destroy the stones, which would also destroy the town's magical water source.

\section{Major NPCs}

\npctitle{Elara (Chorus Leader)}
\begin{itemize}[leftmargin=*]
    \item A once-respected Cantor who discovered the Veridian Stones' power
    \item Now at Tier 5 corruption (Omniscient Gaze)
    \item Believes she is doing the right thing but is becoming increasingly detached from humanity
    \item \textbf{Quirk:} Speaks in third person, referring to herself as "the Chorus"
    \item \textbf{Key Trait:} "I see the truth in all things. It is my duty to share it."
\end{itemize}

\npctitle{Brother Aris (First Truth-Teller)}
\begin{itemize}[leftmargin=*]
    \item The cult's second-in-command
    \item Tier 3 corruption (Revelation's Power)
    \item Was a blacksmith before joining the cult
    \item \textbf{Quirk:} Can see through deception but struggles with his own past lies about his wife's death
    \item \textbf{Key Trait:} "The truth must be spoken, even when it burns."
\end{itemize}

\npctitle{Sister Maren (First Truth-Teller)}
\begin{itemize}[leftmargin=*]
    \item A former midwife
    \item Tier 3 corruption (Revelation's Power)
    \item Has become the cult's healer but is tormented by the truths she's heard during childbirth
    \item \textbf{Quirk:} Carries a small bell that she rings when she senses deception
    \item \textbf{Key Trait:} "The body remembers what the mouth denies."
\end{itemize}

\npctitle{Kaelen (The Captive)}
\begin{itemize}[leftmargin=*]
    \item The missing merchant
    \item Discovered the stones' true danger
    \item Has been imprisoned but not harmed (the cult doesn't believe in physical violence)
    \item \textbf{Quirk:} Has figured out how to temporarily disable the stones' effect
    \item \textbf{Key Trait:} "I was wrong to think truth could be forced."
\end{itemize}

\section{The Adventure}

\subsection{Phase 1: The Truth That Was Spoken}
The PCs arrive at Veridian Crossroads and are immediately stopped at the crossroads. They learn about the cult's truth-enforcing practices and discover Kaelen's last known location.

\keypoint{Key Decisions:}
\begin{itemize}[leftmargin=*]
    \item How will they navigate the truth requirement? (Honesty can be strategic)
    \item Will they try to blend in or confront the cult directly?
    \item Do they believe Kaelen's story about the stones?
\end{itemize}

\subsection{Phase 2: The Truth That Cannot Be Hidden}
The PCs investigate the town and learn about the Veridian Stones. They discover the truth about the stones' effect and why Kaelen wanted to destroy them.

\keypoint{Key Challenges:}
\begin{itemize}[leftmargin=*]
    \item Navigating the Truth Circle without revealing their true intentions
    \item Dealing with the mental strain of the truth-revealing effect
    \item Finding a way to reach Kaelen without the cult's knowledge
\end{itemize}

\subsection{Phase 3: The Truth That Divides}
The PCs discover that the cult is fracturing. Some members want to expand the truth-revealing effect, while others want to destroy the stones.

\keypoint{Key Choices:}
\begin{itemize}[leftmargin=*]
    \item Support the "Purists" who want to expand the truth-revealing effect
    \item Support the "Silencers" who want to destroy the stones
    \item Try to find a middle ground that preserves the town's water source while stopping the mental breakdowns
\end{itemize}

\subsection{Phase 4: The Truth That Must Be Chosen}
The final confrontation at the Chorus House. The PCs must decide:

\keypoint{Critical Choice:}
\begin{itemize}[leftmargin=*]
    \item Free Kaelen and destroy the stones (saving the town but destroying the water source)
    \item Help Elara find a way to control the stones' power (preserving the water but risking further corruption)
    \item Allow the cult to continue their work (saving the water but dooming the town to mental breakdown)
\end{itemize}

\section{Consequences}

\subsection{Success Path}
\begin{itemize}[leftmargin=*]
    \item The PCs help Elara find a way to control the stones' power
    \item Elara becomes the town's guardian, using the stones to reveal only the most critical truths
    \item The cult transforms into the "Veridian Council" - a group that advises the town while respecting necessary privacy
    \item Kaelen stays to help rebuild the town
\end{itemize}

\subsection{Failure Path}
\begin{itemize}[leftmargin=*]
    \item The cult expands its influence, causing mental breakdowns across the region
    \item The town is destroyed when the stones' effect becomes too strong
    \item Kaelen is killed when he tries to destroy the stones
\end{itemize}

\subsection{Antagonistic Path}
\begin{itemize}[leftmargin=*]
    \item The cult becomes a regional threat, forcing neighboring towns to either submit or be destroyed
    \item The PCs must lead a resistance against the Silent Choir's truth-revealing crusade
    \item Elara becomes a tragic villain, fully consumed by her need to share the truth
\end{itemize}

\section{Mechanical Integration}

\begin{itemize}[leftmargin=*]
    \item \textbf{Truth Check Mechanics:} Any time the PCs are in the town, they must make a Spirit + Resolve check (DV 3) or be affected by the truth-revealing effect
    
    \item \textbf{Corruption Spread:} If PCs spend time in the town without leaving, they begin to accumulate corruption (1 segment per day)
    
    \item \textbf{The Chalice of Clarity:} Drinking from it gives +1d to all Truth Circle rolls for 24 hours, but increases corruption risk
\end{itemize}

\section{Why This Works as a Silent Choir Adventure}

\begin{itemize}[leftmargin=*]
    \item \textbf{Thematic Unity:} The entire adventure revolves around the core theme of truth vs. deception
    
    \item \textbf{Corruption Spread:} The cult's structure directly reflects the Silent Choir corruption table
    
    \item \textbf{Moral Complexity:} There's no clear "good" or "bad" side - the cult is both destructive and well-intentioned
    
    \item \textbf{Player Agency:} Multiple paths with meaningful consequences
    
    \item \textbf{Cult Mechanics:} The adventure provides clear ways for the cult to spread its influence (through the Truth Circles and Chalice of Clarity)
\end{itemize}

This adventure demonstrates why The Silent Choir creates the most compelling cults in Fate's Edge - it's not about mind control or domination, but about the most fundamental human dynamic: what happens when we can no longer hide from the truth? The cult isn't a typical villain, but a community that has gone too far in the pursuit of a noble goal, creating a situation where the players must decide what price is too high to pay for truth.

\end{document}