\documentclass[12pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{multicol}

% Define colors
\definecolor{eckpurple}{RGB}{128, 0, 128}
\definecolor{eckgold}{RGB}{218, 165, 32}
\definecolor{eckred}{RGB}{178, 34, 34}
\definecolor{eckmarble}{RGB}{248, 248, 255}

% Section formatting
\titleformat{\section}
{\color{eckpurple}\normalfont\Large\bfseries}
{\color{eckpurple}\thesection}{1em}{}

\titleformat{\subsection}
{\color{eckgold}\normalfont\large\bfseries}
{\color{eckgold}\thesubsection}{1em}{}

% Item spacing
\setlist{itemsep=2pt}

\title{\color{eckpurple} \textbf{The Usurper's Gambit}\\ \large An Adventure for Fate's Edge - Ecktoria}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Adventure Overview}

\textbf{Title:} The Usurper's Gambit\\
\textbf{Region:} Ecktoria -- Marble, Brass, \& Fire\\
\textbf{Theme:} Legitimacy, bureaucracy, ritual power, imperial decline\\
\textbf{Level:} Veteran (Tier III)\\
\textbf{Length:} 4--5 Sessions

\section{Premise}

Ecktoria, heir to the shattered Utaran Empire, stands on the knife-edge of civil conflict. A charismatic pretender has seized the loyalty of the capital militia, silenced dissenting censores, and begun invoking ancient rites to claim the Mandate of Everflame. Meanwhile, the recognized heir --- King Aldric the Lame --- clings to legitimacy through law, tradition, and the blessing of the High Temple.

The players navigate a city where every marble colonnade hides a conspiracy, every mosaic conveys a political message, and every oath is engraved in red-lacquered tablets guarded by the Censors. They must decide: bolster one claimant, forge a third path, or engineer a peaceful transfer of power in a system designed to resist change violently.

\section{Hook}

Choose or roll:

\begin{itemize}
  \item \textbf{The Envoy's Humiliation:} Before the Great Baths of Aurix, the Lame King's Envoy is publicly mocked, their ceremonial staff struck from their hand --- a grave insult in Ecktoria. The crowd cheers the Pretender's heralds.
  \item \textbf{A Dead Man’s Seal:} A courier collapses into the players’ arms with a sealed tablet bearing the mark of a supposedly-dead claimant. If true, a surviving cousin could upend every faction's plans.
  \item \textbf{A Contract Invalidated:} A lucrative patronage contract is voided by a Senate Decree --- half the Senate claims the decree is legal, the other half claims the seals were forged.
\end{itemize}

\section{Key Factions}

\subsection{The Lame King's Line}
Not medieval nobility --- but \textbf{imperial-bureaucratic loyalists}. They adhere to:
\begin{itemize}
  \item \textbf{High Priest of Everflame:} Wielder of ceremonial fire and keeper of coronation rites.
  \item \textbf{Coin-House Factors:} Masters of imperial credit and tax ledgers.
  \item \textbf{Veteran Standard-Bearers:} The disciplined remnant of the old Utaran Legion, clad in lamellar bronze and crimson silk.
\end{itemize}

Their power is \textbf{ritualistic and institutional}. Their weakness: \textit{slow, cautious, inflexible.}

\subsection{The Pretender’s Cause}
Duke Marcus promises a ``New Augustan Revival" and wields:
\begin{itemize}
  \item \textbf{Reformist Senators:} Toga-clad politicians seeking relevance in a fading state.
  \item \textbf{Condotta Captains:} Foreign-trained mercenary cohorts who thrive in urban conflict.
  \item \textbf{The New Merchants:} Silk caravans, spice importers, and canal guilds wanting deregulation.
\end{itemize}

Their power is \textbf{popular and military}. Their weakness: \textit{shallow legal standing.}

\subsection{The Silent Opposition}
They reject both claimants and imagine a return to provincial autonomy.
\begin{itemize}
  \item \textbf{Guild Conspiracies:} Especially the \textit{Marblewrights}, whose control of monuments lets them encode political messages in stone.
  \item \textbf{Shadow Council:} Disaffected nobles who meet in abandoned bathhouses.
  \item \textbf{Regional Governors:} Each with their own armed retinues, quietly waiting.
\end{itemize}

Their power: \textbf{economic sabotage}.  
Their weakness: \textit{disunity.}

\section{Key NPCs}

\subsection{King Aldric the Lame}
A legalist monarch steeped in ceremony, whose limp is from bearing the sacred Everflame during the siege of Ecktoria. Pious, cautious, but deeply principled.

\subsection{Duke Marcus the Bold}
Glossy armor, public charisma, and immaculate timing. He invokes ancient Utaran military rites --- even if he barely understands them.

\subsection{Grand Magistrate Valeria Thorne}
A razor-edged legal scholar who knows every precedent for imperial succession. Her chambers overflow with red tablets inscribed with shame and honor. She alone can interpret the \textit{Pact of Three Flames}, the ancient document defining lawful coronations.

\subsection{High Priest Dominicus}
Keeper of the Sacred Brazier. His blessing determines which claimant the people accept as crowned by the Everflame. Torn between stability and renewal.

\subsection{The Lame King's Envoy}
Masked, elegant, and unsettlingly silent. Rumored to have been a Senator, a spymaster, or Aldric’s illegitimate sibling.

\section{Key Locations}

\subsection{The Grand Forum}
Not a medieval square, but a marble theatre of politics: multi-level balconies, colored awnings, statues of emperors with missing fingers, and mosaic floors depicting the victories of old Utar.

\subsection{Imperial Arena}
Still used for trials by ordeal, military demonstrations, and ceremonial oaths. Beneath it lie tunnels once used for exotic beasts --- now used for political prisoners.

\subsection{Coin-House Floor}
A Byzantine-style financial temple. Gold seals, abacus-boards, clerks chanting interest schedules, and styluses scratching into wax tablets. A single misplaced ledger entry can bankrupt a province.

\subsection{Censor’s Hall}
A labyrinth of documents, rank-lists, shame-tablets, and censored histories. Red lacquer glows like dried blood under the lantern light. The Hall’s archivists maintain \textit{The Line of Fire}, a genealogical record of every legitimate claimant.

Access is tightly controlled and politically explosive.

\subsection{The Fire-Blessed Basilica}
The spiritual heart of Ecktoria. The ceiling is a dome of gold and lapis, depicting the Everflame descending upon the First Emperor. The final coronation ritual must occur here.

\section{Adventure Structure}

\subsection{Act I --- Cracks in Marble}
The PCs witness an event signaling that the Pretender’s forces are seizing soft power: forged decrees, altered statues, or militia patrols wearing new colors. They must investigate whether Duke Marcus’s claim is real or subterfuge.

\subsection{Act II --- Shadows in the Censorium}
The Grand Magistrate is attacked by masked agents attempting to steal the \textit{Pact of Three Flames}. PCs must navigate the Censor's Hall, uncover forged genealogies, or extract truths from censored archives.

Choices here determine which claimant gains legal ground.

\subsection{Act III --- Coin, Steel, and Fire}
The Pretender blockades the Basilica, controlling the square. Loyalist Standard-Bearers muster but cannot act without legal sanction. PCs mediate, sabotage, or escalate:

\begin{itemize}
  \item Reveal a hidden legitimate claimant.
  \item Break the blockade via mercantile leverage in the Coin-House Floor.
  \item Expose Marcus’s forged lineage.
  \item Convince Dominicus to deny Marcus his blessing.
\end{itemize}

\subsection{Act IV --- The Coronation of Ash or Glory}
At dawn, the Basilica doors open. The coronation ritual begins. PCs decide which claimant steps forward --- or whether the old imperial rites are subverted entirely.

The \textbf{Trial of the Everflame} may occur:
\begin{itemize}
  \item A rhetorical duel before the people.
  \item A ritual combat under priestly supervision.
  \item A miraculous display (real or faked).
\end{itemize}

\section{Endings}

\subsection{Aldric Crowned Lawfully}
A stable but static Ecktoria that prizes tradition. PCs may become honored advisors --- or tools of a rigid bureaucracy.

\subsection{Marcus Claims the Throne}
A revitalized but turbulent empire driven by reform. PCs navigate rapid militarization, mercantile upheaval, and foreign ambitions.

\subsection{A Third Path}
PCs install:
\begin{itemize}
  \item The hidden claimant.
  \item A Council of Three Flames.
  \item Regional autonomy under imperial oversight.
\end{itemize}

This ending creates future adventure hooks across Ecktoria.

\subsection{Ecktoria Falls}
If diplomacy, legality, and ritual fail, the city collapses into street warfare.  
The PCs become the only force capable of rebuilding order --- or escaping the ashes.

\end{document}
