\documentclass[10pt,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{a5paper, margin=1.2cm, top=1.5cm, bottom=1.8cm}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{titletoc}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[pdfusetitle]{hyperref}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\tcbuselibrary{skins,breakable}

% Color palette from other Fate's Edge documents
\definecolor{bronze}{RGB}{205,127,50}
\definecolor{iron}{RGB}{70,70,70}
\definecolor{parchment}{RGB}{245,240,225}
\definecolor{blood}{RGB}{120,30,30}
\definecolor{deepblue}{RGB}{20,45,80}
\definecolor{recap}{RGB}{230,220,180}
\definecolor{velvet}{RGB}{120,20,50}
\definecolor{shadow}{RGB}{40,40,60}

% Page style
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE]{\slshape\leftmark}
\fancyhead[RO]{\slshape\rightmark}
\fancyfoot[LE,RO]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% Section formatting
\titleformat{\chapter}[hang]{\Huge\bfseries\color{blood}}{\thechapter}{20pt}{}
\titleformat{\section}{\Large\bfseries\color{iron}}{\thesection}{15pt}{}
\titleformat{\subsection}{\normalsize\bfseries\color{deepblue}}{\thesubsection}{12pt}{}

% Tcolorbox for sidebars
\newtcolorbox{sidebar}[1][]{colback=parchment, colframe=bronze, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{recapbox}[1][]{colback=recap, colframe=bronze!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{mechanicbox}[1][]{colback=gray!10, colframe=iron!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{tipsbox}[1][]{colback=green!5, colframe=green!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{clocksidebar}[1][]{colback=red!5, colframe=red!70, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{heistbox}[1][]{colback=velvet!5, colframe=velvet!80, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}
\newtcolorbox{elegybox}[1][]{colback=shadow!10, colframe=shadow!80, sharp corners, breakable, enhanced, fonttitle=\bfseries, title=#1}

% Custom commands for mechanics
\newcommand{\tag}[1]{\textsc{[#1]}}
\newcommand{\sb}{\textsc{sb}}
\newcommand{\dv}{\textsc{dv}}
\newcommand{\xp}{\textsc{xp}}

\begin{document}

\title{\Huge The Velvet War of Silkstrand \\[0.5cm] \large A Prequel Adventure for Fate's Edge}
\author{The last testament of the Silk Coin}
\date{Year of the Falling Bridge}

\maketitle
\thispagestyle{empty}

\chapter*{How to Run This Prequel}
\addcontentsline{toc}{chapter}{How to Run This Prequel}

\begin{elegybox}[A Note on Tragedy]
This is not a standard adventure. The player characters are \textbf{doomed}. They will not survive the final act. Their goal is not victory but \textit{delay} — to buy time for the pre‑generated exiles of \textit{Blood and Silk} to escape. Embrace the tragedy. Let each death be a choice, each sacrifice a gift to the future.

The mechanics support this: scripted death scenes, Legacy Tokens, and a visible \textbf{Court's Net Clock} that ticks inexorably toward the guildhouse's fall.
\end{elegybox}

\textit{The Velvet War of Silkstrand} is a Tier II–III prequel adventure, designed for 3–5 players who will create original members of the \textbf{Silk Coin}, a small, principled thieves' guild. The pre‑generated characters from \textit{Blood and Silk} (Kestra, Marcus, Sariel, Elena, Thane, Lyra, Finn, Zara) appear as junior members — apprentices who witness the heroes' final stand.

\begin{recapbox}[What the Players Should Know]
\begin{itemize}
\item The Silk Coin was once a modest guild of fence, smugglers, and honest thieves.
\item Their patron is \textbf{Old Kes}, a fence who taught them the trade.
\item The \textbf{Velvet Court} is a ruthless syndicate that now demands the Silk Coin's territory.
\item A mysterious power — tied to a brothel called \textbf{The Gilded Pillow} — is turning the Court's enemies into willing collaborators.
\end{itemize}
\end{recapbox}

\section{The Tragedy's Shape}

The adventure unfolds in three acts:

\begin{enumerate}
\item \textbf{Act I – The Silk Coin's Last Score:} The guild attempts a heist to buy Old Kes's freedom. It is a trap.
\item \textbf{Act II – The Hunt:} The Velvet Court systematically destroys the Silk Coin's allies, resources, and safe houses.
\item \textbf{Act III – The Burning Guildhouse:} The final assault. The pre‑gens escape through a hidden tunnel while the player characters hold the line.
\end{enumerate}

Player characters \textbf{will die}. The only question is how many of the eight pre‑gens survive — and what Legacy Tokens they carry into \textit{Blood and Silk}.

\section{New Mechanics}

\subsection{Desperation Dice}

When a player character is at \textbf{Harm 2} (their second wound), they may roll an extra d10 on any action. On a roll of 1–9, the action proceeds normally. On a \textbf{10}, they succeed spectacularly but immediately take their third wound (incapacitated). This represents a last, heroic burst.

\subsection{Court's Net Clock (Visible)}

\begin{longtable}{p{4cm} c p{5cm}}
\toprule
\textbf{Clock Segments} & \textbf{Effect} \\
\midrule
0–2 & The Court is probing. Safe houses are still hidden. \\
3–4 & A contact is turned. One Safe House is lost. \\
5–6 & The guildhouse is surrounded. No one enters or leaves without being seen. \\
7 & Reinforcements arrive (Black Banner mercenaries). \\
8 & The final assault begins. \\
\bottomrule
\end{longtable}

The clock advances when:
\begin{itemize}
\item The party fails a Stealth or Subterfuge roll (+1).
\item The GM spends 2 \sb to represent the Court's intelligence network (+1).
\item A player character is captured or killed (+2).
\item The party lingers in one location for too long (GM's discretion, +1).
\end{itemize}

\subsection{Legacy Tokens}

At the end of the adventure, the GM awards \textbf{Legacy Tokens} based on how many pre‑gens escape:

\begin{itemize}
\item \textbf{8 escape (heroic):} 3 Legacy Tokens.
\item \textbf{6–7 escape:} 2 Legacy Tokens.
\item \textbf{Fewer than 6:} 1 Legacy Token.
\end{itemize}

Each token becomes a \textbf{minor Asset} in \textit{Blood and Silk}, chosen by the players who played the prequel. Examples:

\begin{itemize}
\item \textit{Elena's Medicine Stash} – A hidden cache of rare herbs (+1 die to Medicine rolls once).
\item \textit{Marcus's Worn Sword} – A blade that never dulls (ignore one \textit{Compromised} tag).
\item \textit{Sariel's Lockpick Set} – +1 die to open any lock that was touched by a Silk Coin member.
\item \textit{Thane's Toolkit} – Can repair one piece of \textit{Compromised} gear without a roll.
\item \textit{Lyra's Diplomatic Papers} – A forged letter of introduction to a Vilikari lord.
\item \textit{Finn's Fox Companion} – A Cap 1 follower that can sniff out danger.
\item \textit{Zara's Ledger} – A list of Velvet Court contacts (grants +1 die to Streetwise in Silkstrand).
\item \textit{Kestra's Scroll} – A one‑use Rite of the Traveler (any low rite).
\end{itemize}

\section{Factions and Key NPCs}

\subsection{The Silk Coin (Player Characters)}
A small guild of 3–5 thieves, each with a specialty. Players should create their own characters using the \textit{Fate's Edge} core rules (Tier II, 50–70 \xp). Suggested archetypes:

\begin{itemize}
\item \textbf{The Mastermind} (Wits, Subterfuge, Command) – plans the heists.
\item \textbf{The Ghost} (Wits, Stealth, Ikasha rites) – infiltration.
\item \textbf{The Blade} (Body, Melee, Athletics) – muscle.
\item \textbf{The Silver Tongue} (Presence, Sway, Livaea rites) – social infiltration.
\item \textbf{The Tinker} (Wits, Craft, Tinker) – traps and devices.
\end{itemize}

\subsection{The Pre‑Gens (Non‑Player Characters)}
These eight are junior members, not yet full thieves. They have Cap 1 skills and no Talents. Their role is to be \textit{protected} and \textit{evacuated}. The GM should describe their fear, their hope, and their growing admiration for the player characters.

\subsection{The Velvet Court}

\textbf{Madam Serafine} (Tier IV, off‑screen) – The Court's true ruler. She is never seen directly, but her influence touches every scene. She speaks through intermediaries, always cloaked, her face hidden by a velvet veil. Her voice is warm, reasonable, and utterly without mercy.

\textbf{The Velvet Enforcer (Antagonist)} – A Cap 3 opponent who leads the attacks. She wears a silver mask and fights with a poisoned stiletto. Her name is never revealed; the Court calls her \textit{the Silken Blade}.

\textbf{The Gilded Pillow (Livaean Brothel)} – A front for the Court's intelligence network. The courtesans are Livaean cultists who extract secrets from clients. Entering the Pillow with hostile intent triggers a \tag{WARD} (the goddess protects her own). The players may discover that Old Kes's betrayal was compelled by a Livaean \textit{Unrefusable Offer} he could not refuse.

\subsection{Old Kes (Terrestrial Patron, Betrayer)}
The Silk Coin's mentor and fence. He is captured in Act I, but the twist in Act III reveals he was never captured — he \textit{offered} himself as bait. Livaea compelled him with an oath he could not break. His betrayal is not malice; it is desperation. He loved the Silk Coin. He also loved his daughter, who is held hostage at the Gilded Pillow.

\subsection{Mareth (Secret Ally)}
The hooded scholar from \textit{Blood and Silk} appears in Act III, guiding the pre‑gens through the canal tunnels. She does not fight. She only whispers: \textit{"This way. I knew your teachers. They asked me to keep you safe."} She does not explain how she knew.

\chapter{Act I — The Silk Coin's Last Score}

\section{Opening Scene — Old Kes Is Taken}

The adventure begins in the Silk Coin's guildhouse, a converted spice warehouse near the Dye Yards. Old Kes was supposed to return at dawn. He did not. A terrified messenger — a street urchin the guild protects — arrives with a bloodied note: \textit{"The Bronze Ledger. One night. Or Kes hangs."}

The Bronze Ledger is a blackmail book kept by Madam Serafine herself. It contains secrets of every major official in Silkstrand. Stealing it is suicide.

But the Silk Coin has no choice.

\section{Scene 1 — The Guildhouse}

Introduce the player characters and the pre‑gens. Ask each player to describe their character's relationship with one pre‑gen (e.g., \textit{"I taught Kestra how to pick a lock."} or \textit{"Marcus once saved my life from a Watch patrol."}). These relationships become Bonds that generate Boons later.

The pre‑gens are:
\begin{itemize}
\item \textbf{Kestra} — copying scrolls in a corner, her fingers stained with ink.
\item \textbf{Marcus} — sharpening a sword, brooding.
\item \textbf{Sariel} — perched on a windowsill, watching the street.
\item \textbf{Elena} — tending a wounded urchin.
\item \textbf{Thane} — working a forge, muttering about ventilation.
\item \textbf{Lyra} — reading a letter from her family, then hiding it.
\item \textbf{Finn} — stroking a grey fox with a missing ear.
\item \textbf{Zara} — counting silver, her face pale.
\end{itemize}

\subsection{The Bronze Ledger — Planning the Heist}

The players have one night to case the Velvet Court's headquarters, a mansion in the Patrician Heights. The GM should run this as a \textbf{heist planning montage} using the mechanics from \textit{The Velvet Touch}. Each player describes one preparation (casing the guard patrol, bribing a servant, mapping the sewers). Each strong success on a Setup roll (Wits+Stealth or Wits+Streetwise, \dv 3) grants +1 Position or –1 \dv later.

The players learn that the Bronze Ledger is kept in a lead‑lined safe in Serafine's private study. The study is guarded by a \tag{WARD} inscribed with Livaea's sigil — a kiss and a thorn.

\subsection{The Heist — A Trap}

Run the heist as a series of challenges. The players succeed in reaching the safe — but the ledger is not there. Instead, they find a single sheet of parchment: \textit{"Welcome. The Court has watched you for a year. Tonight, we collect."}

The mansion's doors lock. The \tag{WARD} turns hostile (burns any non‑Livaean for 1 Fatigue per round). The players must escape. Use the pursuit rules from \textit{Violets and Stone} (Chase through Patrician Heights, DV 3–5). On success, they return to the guildhouse with a few minor ledgers (enough to buy a day). On failure, one player character is captured (they will be rescued in Act II).

\textbf{End of Act I — Court's Net Clock advances to 2.}

\chapter{Act II — The Hunt}

\section{Opening Scene — The Pillow House's Whisper}

A contact tells the Silk Coin that Old Kes was seen entering the \textbf{Gilded Pillow}, a Livaean brothel, the night before his disappearance. The Pillow House is neutral ground — no violence is permitted within its walls. But secrets are not.

The players may infiltrate the Pillow House using social skills (Presence+Sway, Presence+Performance). A successful roll reveals that Kes was not imprisoned — he \textit{volunteered}. A courtesan (a Livaean priestess) confesses under pressure: \textit{"His daughter is here. She serves the goddess now. Kes would do anything to free her."}

If the players investigate further, they uncover that the \textit{Unrefusable Offer} was used on Kes. He cannot disobey. His betrayal is a tragedy, not a treachery.

\subsection{The Court's Retaliation}

The Velvet Court strikes back. Use a series of short, brutal vignettes:

\begin{itemize}
\item \textbf{The Fence is Taken:} The Silk Coin's secondary fence is arrested. The Watch arrives with a warrant signed by a Censor who owes the Court.
\item \textbf{The Safe House Burns:} A fire in the Dye Yards destroys a hidden cache of supplies. The players may rescue the pre‑gens from the flames.
\item \textbf{The Informant is Turned:} A trusted ally sells the guild's patrol schedules. The Court's assassins ambush a player character on a lonely bridge.
\end{itemize}

Each vignette ticks the \textbf{Court's Net Clock} by +1. The GM should let the players attempt to mitigate the damage (e.g., bribing the Censor, fighting the assassins) but the clock \textit{will} advance.

\subsection{The Betrayal Revealed}

Old Kes meets the players in secret — in the catacombs beneath the Gilded Pillow. He is gaunt, his eyes hollow. He confesses everything: Livaea's compulsion, his daughter's captivity, his hopelessness.

He offers one last gift: the combination to the Bronze Ledger's backup safe, hidden in his own warehouse. He also warns them: \textit{"The Court's net closes at dawn. Get the children out."}

If the players spare him, he helps them escape. If they condemn him, he accepts his fate. Either way, he does not survive the night.

\textbf{End of Act II — Court's Net Clock at 5 or 6.}

\chapter{Act III — The Burning Guildhouse}

\section{Opening Scene — The Night Before}

The guildhouse is surrounded. The players have one hour to prepare. They can:

\begin{itemize}
\item \textbf{Barricade the doors} (Body+Craft, \dv 3) – ticks the final assault back by 1 segment.
\item \textbf{Send the pre‑gens to the canal tunnel} (Wits+Stealth, \dv 4) – buys time for the escape.
\item \textbf{Negotiate with the Court} (Presence+Sway, \dv 5) – a false offer; the Court will betray them.
\item \textbf{Write a letter} – each player may write a short note to a pre‑gen. This becomes a Legacy Token.
\end{itemize}

\subsection{The Final Assault}

The Court sends its best: the \textbf{Silken Blade} (Cap 3, Harm 3, Stealth 4, can poison) and a squad of Black Banner mercenaries (Cap 2 squad). The players must hold the entrance while the pre‑gens escape through the tunnel.

Run this as a \textbf{scripted combat}. The players cannot win. Instead, they choose how to spend their remaining resources:

\begin{itemize}
\item \textbf{Spend a Boon} to delay the Court for one round.
\item \textbf{Sacrifice a Limb} – permanently reduce an Attribute by 1 to save two pre‑gens.
\item \textbf{Use a Desperation Die} – as described above. A 10 means a spectacular last stand.
\end{itemize}

\subsection{Mareth's Arrival}

Midway through the battle, a hooded figure appears at the canal entrance — \textbf{Mareth}. They do not fight. They simply wave the pre‑gens forward. \textit{"This way. I knew your teachers. They asked me to keep you safe."}

The pre‑gens hesitate. The player characters shout at them to run. They run.

\subsection{The Last Stand — Player Character Deaths}

Each player character gets a \textbf{final scene}. Describe their death in as much detail as they wish. The GM may ask: \textit{"What is the last thing you see?"} or \textit{"What do you say to the children as they flee?"}

When all player characters are down, the GM describes the guildhouse collapsing. The Court's enforcers walk through the flames, searching for survivors. They find none of the pre‑gens.

\subsection{Epilogue — The Canal}

The pre‑gens emerge from a storm drain on the river's edge. They are ragged, silent. Mareth is gone. Kestra finds a note pinned to a post: \textit{"Go west. Do not look back. The road will teach you."}

They look back. The sky above the Dye Yards is red.

One of them — Marcus, perhaps — speaks: \textit{"They bought us time. We'll not waste it."}

They walk.

\section*{Awarding Legacy Tokens}

At the end of the session, the GM awards Legacy Tokens based on how many pre‑gens escaped (all eight, by default, unless the players chose to sacrifice some). Each player chooses one Legacy Asset to appear in \textit{Blood and Silk}. The GM notes these for the campaign.

\chapter{Appendix — NPC Quick Stats}

\subsection{The Silken Blade (Velvet Court Enforcer)}
Tier III. Body 3, Melee 4, Stealth 4. Talents: \textit{Poisoned Blade} (on a hit, target tests Resolve \dv 3 or gains 1 Fatigue), \textit{Shadow Meld} (Ikasha's rite), \textit{The Velvet Glove} (can call a minor NPC as a distraction). Harm 3.

\subsection{Old Kes (Broken Fence)}
Tier II. No combat. Offers Strings (\textit{Kes's Warehouse Key}, \textit{Kes's Last Ledger}). Under Livaea's compulsion; cannot resist a direct order from a Courtesan.

\subsection{Livaean Courtesan (Gilded Pillow)}
Tier II. Presence 3, Sway 3, Performance 2. Rites: \textit{Golden Tongue}, \textit{The Unrefusable Offer} (once). Will not fight; will try to seduce or bargain.

\subsection{Mareth (Mysterious Ally)}
No combat. Wits 4, Lore 4, Traveler's rites. Does not speak of their past. Appears only to guide the pre‑gens.

\subsection{Black Banner Mercenary (Squad)}
Cap 2 squad. Body 3, Melee 2, Athletics 2. Harm 2 per squad member. Two squads are present in the final assault.

\chapter{Colophon}

This account was written by the last surviving clerk of the Silk Coin, on paper salvaged from the burning guildhouse. The ink is lamp‑black and ash. The binding is a torn ledger cover.

The coin was spent. The debt was paid. The road remembers.

\end{document}