\documentclass[12pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{mdframed}
\usepackage{needspace}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{calc}
\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{longtable}
\usepackage{pdflscape}
\usepackage{multicol}
\usepackage{graphicx}

\definecolor{darkblue}{rgb}{0.0, 0.0, 0.55}
\hypersetup{
    colorlinks=true,
    linkcolor=darkblue,
    urlcolor=darkblue,
    pdftitle={The Vow of Broken Glass},
    pdfauthor={Fate's Edge Adventure Module}
}

% ------------------- Custom commands -------------------
\newcommand{\tag}[1]{\textsc{#1}}
\newcommand{\clk}[1]{\textit{[#1]}}
\newcommand{\stringitem}[1]{\textit{#1}}
\newcommand{\dusana}[1]{%
    \begin{quote}\itshape #1 ---~Dusana of the Raven Road\end{quote}}
\newcommand{\letter}[2]{%
    \begin{quote}\small\itshape #1 \hfill --- #2\end{quote}}

% ------------------- Styling -------------------
\titleformat{\chapter}{\Huge\bfseries}{\thechapter}{20pt}{}
\titleformat{\section}{\Large\bfseries}{\thesection}{15pt}{}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8ex}

% ------------------- Infobox styling -------------------
\newmdenv[
    linecolor=black,
    backgroundcolor=gray!5,
    innertopmargin=8pt,
    innerbottommargin=8pt,
    innerrightmargin=8pt,
    innerleftmargin=8pt,
    skipabove=6pt,
    skipbelow=6pt,
    font=\small
]{infobox}

% ------------------- Document -------------------
\begin{document}

\title{
    \Huge\textbf{The Vow of Broken Glass}\\[0.5em]
    \Large A Fate's Edge Adventure
}
\author{
    Tier II--III (2--3 Sessions)\\
    \small Revenge Fantasy Series
}
\date{
    \small Written in blood and ink
}
\maketitle

\tableofcontents

\chapter{Overview: The Economy of Vengeance}

Revenge is a dish best served with proper documentation. Every wound has a receipt. Every grave has a marker. And every oath has a clause that someone will eventually exploit.

\vspace{1em}
\dusana{I once knew a man who spent twenty years sharpening a single knife. He never used it. He died of old age holding the edge. The knife was sharp enough to cut memory. I keep it on my shelf. It reminds me: preparation is not action.}

\section{Premise}

The party is hired by \textbf{Lady Aria Vane}, a noblewoman whose family was massacred by a rogue faction of the \textbf{Iron Syndicate}. She provides a list of five names. Each name corresponds to a lieutenant responsible for the attack. The party must hunt them down, extract information, and eliminate them. However, each kill advances the \textbf{Blood Debt Clock}. When full, the Syndicate retaliates against something the party loves. The twist: Lady Vane is not who she claims to be. She is the architect of the conflict, using the party to clean up loose ends from a failed coup.

\section{Themes}

\begin{itemize}
    \item \textbf{The Cost of Vengeance:} Every action requires a sacrifice.
    \item \textbf{Bureaucracy of Violence:} Killing requires planning, supplies, and cover-ups.
    \item \textbf{The Cycle:} Violence begets violence; the hunter becomes the hunted.
    \item \textbf{Homage/Parody:} Treats revenge tropes (the list, the training, the final duel) with deadly seriousness, highlighting the absurdity of the cost.
\end{itemize}

\section{Tone}

Gritty, methodical, intense. The humor comes from the juxtaposition of high-stakes violence with mundane logistical hurdles (e.g., \textit{I cannot duel the warlord today; I must file a death certificate to avoid scrutiny}). Play the tropes straight. The absurdity is in the world, not the characters.

\section{Key Mechanics}

\begin{infobox}
\textbf{Adventure Mechanics}
\begin{itemize}
    \item \textbf{Blood Debt Clock [6]:} Tracks Syndicate attention. Advances on every kill or major disturbance. When full, retaliation occurs.
    \item \textbf{Vow Tokens:} Each PC pledges a personal sacrifice for the contract (e.g., \textit{I will not sleep until the third name falls}). Spending a token grants +1 Die but marks 1 Fatigue.
    \item \textbf{The List:} A physical handout with five names. Crossing one off triggers a Syndicate Move.
    \item \textbf{Position \& Effect:} Standard SRD rules. Revenge actions often start at \textbf{Desperate} Position due to enemy preparation.
\end{itemize}
\end{infobox}

\section{Locations}

\begin{itemize}
    \item \textbf{The Velvet Room:} Lady Vane's private study. Safe, quiet, filled with portraits of the dead.
    \item \textbf{The Iron Foundry:} Headquarters of the first target. Noisy, hot, guarded.
    \item \textbf{The Whisper Market:} Black market for information and unregistered weapons.
    \item \textbf{The Glass Garden:} Final confrontation site. A greenhouse filled with broken mirrors.
\end{itemize}

\section{Key NPCs}

\begin{itemize}
    \item \textbf{Lady Aria Vane:} The Patron. Elegant, cold, grieving. Secretly the mastermind.
    \item \textbf{Kaelen the Butcher:} Target \#1. Brute force. Loves heavy weapons.
    \item \textbf{Sister Morwenna:} Target \#3. A healer who poisoned the family. Uses Spirit magic.
    \item \textbf{The Syndicate Head:} Target \#5. The true villain (or so it seems).
    \item \textbf{The Auditor:} A neutral figure who tracks deaths for the city guard. Can be bribed or bypassed.
\end{itemize}

\section{The Hook}

The party is summoned to the Velvet Room. Lady Vane offers 5,000 gold marks and a \textbf{Symbol of Patronage} (Invoker access). She slides the List across the table. \textit{Five names. Five deaths. Bring me their rings.}

\chapter{Act I: The Contract}

\section{Scene 1: The Velvet Room}

Lady Vane outlines the terms. The party must agree to the \textbf{Vow of Broken Glass}. Each PC must state one thing they will sacrifice for this job (e.g., \textit{My reputation, my safety, a cherished item}). This creates a \textbf{Vow Token}.

\textbf{Mechanic:} Creating a Vow Token grants 1 Boon. Spending it later grants +1 Die on a revenge-related roll but marks 1 Fatigue.

\letter{They say revenge is a river. I say it is a ledger. Every drop of blood spilled must be accounted for. I am hiring you to balance the books.}{--- Lady Aria Vane}

\section{Scene 2: The First Name (Kaelen)}

Kaelen resides at the Iron Foundry. The party must infiltrate, locate, and eliminate him.

\begin{itemize}
    \item \textbf{Infiltration (Stealth/Subterfuge):} DV 3 (Controlled). On a Miss, advance Blood Debt Clock by 1.
    \item \textbf{Confrontation (Combat):} Kaelen uses Heavy Weapons (Slow Tempo). Position starts at Controlled.
    \item \textbf{Extraction:} Retrieve his ring. This requires an Interact action (DV 2).
\end{itemize}

\textbf{Complication:} The Auditor arrives mid-scene. The party must hide the body or bribe him (Sway DV 4). Failure advances Blood Debt Clock.

\section{Scene 3: The Whisper Market}

To find the next names, the party must visit the market. Information costs more than gold.

\textbf{Currency:} Secrets. The party must trade a secret (GM chooses a minor backstory element) to learn the location of Target \#2.

\textbf{Mechanic:} Trading a secret grants a \textbf{Clue Token}. Spending it reduces DV by 1 for the next hunt.

\chapter{Act II: The Trail}

\section{Scene 4: Targets \#2 and \#3}

The party hunts the remaining lieutenants. This act is a montage of hunts.

\begin{infobox}
\textbf{Montage Mechanics}
\begin{itemize}
    \item Each hunt is a single roll (Wits + Survival or Stealth). DV 4.
    \item \textbf{Success:} Target eliminated. Cross name off List.
    \item \textbf{Partial:} Target eliminated, but Blood Debt Clock advances by 1.
    \item \textbf{Miss:} Target escapes. Blood Debt Clock advances by 2. Party gains 2 Boons.
\end{itemize}
\end{infobox}

\textbf{Revelation:} During the hunt for Sister Morwenna, the party finds a letter in her quarters. It implicates Lady Vane in the original massacre. The Syndicate was hired \textit{by her} to kill her rivals, then she turned on them to silence witnesses.

\section{Scene 5: The Choice}

The party confronts Lady Vane with the evidence. She does not deny it. \textit{Does it matter? They are dead. You are paid. Finish the list.}

\textbf{Decision Point:}
\begin{itemize}
    \item \textbf{Continue:} Finish the contract. Gain full payment. Blood Debt Clock fills automatically at the end.
    \item \textbf{Turn:} Betray Vane. The Syndicate remnants ally with the party against her. Payment is lost.
    \item \textbf{Walk:} Leave the city. Keep payment. Vane hires others to finish the job.
\end{itemize}

\chapter{Act III: The Final Name}

\section{Scene 6: The Glass Garden}

The final confrontation takes place in Vane's private greenhouse. The floor is lined with mirrors (hazardous terrain).

\textbf{Terrain Effect:} Moving through the garden requires Athletics DV 3. Failure causes 1 Harm (shards).

\textbf{Lady Vane's Stats:}
\begin{itemize}
    \item \textbf{Attributes:} Presence 4, Wits 3, Spirit 3.
    \item \textbf{Skills:} Command 3, Arcana 2, Sway 4.
    \item \textbf{Talents:} \textit{Iron Will} (Resist Fear), \textit{Patron's Symbol} (Invoker Rites).
    \item \textbf{Defense:} She uses \textbf{Glass Wards} (Barrier Tags). DV 4 to bypass.
\end{itemize}

\section{Scene 7: The Duel}

This is a social and physical contest. The party must break her wards while resisting her commands.

\begin{itemize}
    \item \textbf{Phase 1 (Wards):} Disable the Glass Wards (Arcana/Tinker DV 4).
    \item \textbf{Phase 2 (Will):} Resist her Command Rites (Spirit + Resolve DV 5).
    \item \textbf{Phase 3 (Strike):} Deliver the final blow (Melee/Ranged DV 3).
\end{itemize}

\textbf{Mechanic:} If the Blood Debt Clock is full, the Syndicate arrives during Phase 3. The party must fight both Vane and the Syndicate (Desperate Position).

\section{Scene 8: The Aftermath}

Vane falls. The List is complete. The party stands amidst the broken glass.

\textbf{Epilogue Options:}
\begin{itemize}
    \item \textbf{The New Masters:} The party takes over Vane's network. They become the new Syndicate.
    \item \textbf{The Burn:} The party burns the Velvet Room. They leave the city, hunted but free.
    \item \textbf{The Ledger:} The party hands the evidence to the City Guard. Vane is arrested. The party is paid nothing but gains \textbf{Reputation}.
\end{itemize}

\chapter{Clocks and Tokens}

\section{Blood Debt Clock [6]}

Advances on every kill, major disturbance, or failed stealth roll.

\begin{infobox}
\textbf{Clock Consequences}
\begin{itemize}
    \item \textbf{Segments 1--3:} Minor scrutiny (guards ask questions, prices rise).
    \item \textbf{Segments 4--5:} Active pursuit (bounties posted, assassins sent).
    \item \textbf{Segment 6:} Retaliation (Syndicate attacks a safe house or ally).
\end{itemize}
\end{infobox}

\section{Vow Tokens}

Created during Character Creation in Act I.

\begin{itemize}
    \item \textbf{Creation:} State a sacrifice. Gain 1 Boon.
    \item \textbf{Spend:} +1 Die on revenge roll. Mark 1 Fatigue.
    \item \textbf{Break:} If the party abandons the vow, they lose all held Boons.
\end{itemize}

\section{Clue Tokens}

Earned in the Whisper Market.

\begin{itemize}
    \item \textbf{Spend:} Reduce DV by 1 for a hunt.
    \item \textbf{Trade:} Give to an NPC for assistance.
\end{itemize}

\chapter{NPC Profiles}

\begin{longtable}{p{0.25\textwidth} p{0.7\textwidth}}
\toprule
\textbf{Lady Aria Vane} & \textbf{Role:} Patron, Antagonist \\
& \textbf{Stats:} Presence 4, Wits 3, Spirit 3 \\
& \textbf{Goal:} Cleanse her history of witnesses. \\
& \textbf{Quote:} \textit{Does it matter? They are dead. You are paid. Finish the list.} \\
\midrule
\textbf{Kaelen the Butcher} & \textbf{Role:} Target \#1 \\
& \textbf{Stats:} Body 4, Melee 3 \\
& \textbf{Goal:} Survive the hunt. \\
& \textbf{Quote:} \textit{Come then. Let us see who bleeds first.} \\
\midrule
\textbf{The Auditor} & \textbf{Role:} Obstacle \\
& \textbf{Stats:} Wits 3, Sway 2 \\
& \textbf{Goal:} Balance the city's death ledger. \\
& \textbf{Quote:} \textit{Every corpse requires a form. Do you have the form?} \\
\midrule
\textbf{Syndicate Enforcers} & \textbf{Role:} Minions \\
& \textbf{Stats:} Body 2, Melee 2 \\
& \textbf{Goal:} Protect the organization. \\
& \textbf{Quote:} \textit{The debt is called. Pay in blood.} \\
\bottomrule
\end{longtable}

\chapter{GM Guidance}

\section{Play the Tropes Straight}

Do not wink at the camera. When a PC monologues about vengeance, treat it as a \textbf{Command Rite}. When they sharpen a blade, treat it as a \textbf{Ritual}. The humor comes from the world reacting logically to extreme behavior.

\section{Manage the Clock}

The Blood Debt Clock is the pressure valve. If the party is too cautious, tick it anyway (rumors spread). If they are too reckless, tick it twice. Ensure it fills by Act III.

\section{The Twist}

Reveal Vane's betrayal in Act II, not Act III. This gives the party time to plan their response. If they wait until the final duel, it feels like a cheap trick. If they know early, it becomes a strategic choice.

\section{Vow Tokens}

Encourage players to make meaningful sacrifices. \textit{I will not drink wine} is weak. \textit{I will not speak until the job is done} is strong. Mechanically reward the stronger vows with narrative weight.

\section{Combat as Drama}

Use the SRD Position rules. A duel in a narrow hallway is \textbf{Desperate}. A ambush from the shadows is \textbf{Dominant}. Make the terrain matter. The Glass Garden should feel hazardous.

\section{Resolution}

There is no \textit{good} ending. Only consequences.
\begin{itemize}
    \item \textbf{Victory:} They kill Vane. They are now the most wanted in the city.
    \item \textbf{Justice:} They arrest Vane. They are broke and owe favors to the Guard.
    \item \textbf{Mercy:} They spare Vane. She hires them again next season.
\end{itemize}

\chapter{Handouts and Props}

\section{The List}

A physical card with five names. Each name has a checkbox.
\begin{enumerate}
    \item Kaelen the Butcher [ ]
    \item Mistress Silas [ ]
    \item Sister Morwenna [ ]
    \item Lord Verrick [ ]
    \item The Syndicate Head [ ]
\end{enumerate}

\section{Vow Cards}

Blank cards for players to write their sacrifices.
\textit{I, [Name], vow to [Sacrifice] until [Condition] is met.}

\section{Syndicate Mark}

A token given when the Blood Debt Clock fills.
\textit{You are marked. Safe houses are no longer safe.}

\chapter{Colophon}

The Velvet Room is empty now. The mirrors are swept. The List is burned. Lady Vane's ring sits in a vault beneath the City Guard, evidence of a crime that never happened.

\vspace{1em}
\dusana{I asked the survivor what he felt when the last name fell. He said, \textit{Lighter.} I asked, \textit{Than before?} He said, \textit{Than alive.} I poured him a drink. He did not take it. He was already full.}

\vspace{1em}
\letter{On the road to the Glass Garden, where the reflections do not match.}{--- Dusana of the Raven Road}

\end{document}
