\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{fancybox}
\usepackage{framed}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{changepage}
\usepackage{mdframed}
\usepackage{sectsty}
\usepackage{tocloft}
\usepackage{etoolbox}
\usepackage{needspace}
\usepackage{parskip}
\usepackage{fontawesome}

\definecolor{shadecolor}{gray}{0.95}
\definecolor{headercolor}{RGB}{30, 50, 100}
\definecolor{accentcolor}{RGB}{70, 130, 180}
\definecolor{tableheader}{RGB}{240, 240, 240}
\definecolor{darkred}{RGB}{139, 0, 0}

\pagestyle{fancy}
\fancyhf{}
\rhead{\thepage}
\lhead{Fate's Edge: Whispers in the Stacks}
\renewcommand{\headrulewidth}{0.4pt}

\sectionfont{\color{headercolor}\normalfont\fontsize{14}{16}\selectfont}
\subsectionfont{\color{accentcolor}\normalfont\fontsize{12}{14}\selectfont}
\subsubsectionfont{\normalfont\fontsize{11}{13}\selectfont}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\geometry{margin=1in}

\newenvironment{monsterentry}[1]{%
  \begin{mdframed}[backgroundcolor=shadecolor, linewidth=0pt, leftmargin=0pt, rightmargin=0pt]%
  \subsection*{#1}%
}{%
  \end{mdframed}%
}

\newenvironment{mechanic}[1]{%
  \begin{mdframed}[backgroundcolor=tableheader, linewidth=1pt, linecolor=accentcolor]%
  \subsubsection*{#1}%
}{%
  \end{mdframed}%
}

\newcommand{\dice}[1]{\texttt{#1}}

\hypersetup{
    colorlinks=true,
    linkcolor=accentcolor,
    filecolor=magenta,      
    urlcolor=accentcolor,
}

\begin{document}

\begin{titlepage}
\centering
\vspace*{2cm}

{\Huge\bfseries\color{headercolor} Whispers in the Stacks} 

\vspace{0.5cm}

{\Large\itshape A Horror Adventure Module for Fate's Edge}

\vspace{2cm}

\vspace{2cm}

{\Large\bfseries Module Type: Horror Adventure}

\vspace{1cm}

{\large Designed for 3-6 players, Tier II-III characters}

\vspace{1cm}

{\large Game Master's Guide Included}

\vfill

{\large 
\textbf{Using the Deck System and Thepyrgos Setting} \\
\textbf{Featuring Entities from Corrupted Scholars to the Whispered Judge} \\
\textbf{Incorporating Dread Clocks and Boon-Based Horror Mechanics}
}

\end{titlepage}

\newpage

\tableofcontents

\newpage

\section{Introduction}

\subsection{Welcome to the University of Thepyrgos}

In the ancient city of Thepyrgos, where knowledge is both currency and curse, the PCs find themselves drawn to the University's towering spires and labyrinthine stacks. What begins as scholarly pursuit quickly transforms into a descent into paranoia and cosmic horror as they discover that the very architecture of learning harbors dark secrets.

This adventure module, "Whispers in the Stacks," challenges your players with unknown threats that lurk in the spaces between knowledge, where isolation breeds fear and the pursuit of truth becomes a dangerous obsession.

\subsection{Module Overview}

\textbf{Adventure Hook:} The PCs are scholars, researchers, or investigators who have been drawn to the ancient University of Thepyrgos, renowned for its vast archives and the mysterious Synod Hall where "judgment is audible at a whisper."

\textbf{Setting:} The University of Thepyrgos - ancient towers, labyrinthine stacks, and the ominous Synod Hall where legal proceedings blend with supernatural influence.

\textbf{Themes:} Isolation, Unknown Threats, Psychological Decay, Escalating Tension, Moral Ambiguity

\textbf{Tone:} Ominous, claustrophobic, and deeply unsettling. Knowledge becomes dangerous, and what seems like scholarly pursuit reveals itself to be a descent into madness.

\textbf{Recommended Character Tier:} Seasoned to Veteran (41-150 XP)

\textbf{Estimated Play Time:} 3 sessions

\section{The University of Thepyrgos - Setting the Scene}

\subsection{What Lies Within These Walls}

The University of Thepyrgos stands as a monument to human knowledge, its towers reaching skyward like fingers grasping for understanding through an uncaring void. But within its ancient stones, something has awakened in the spaces between knowledge — perhaps connected to the "General watch" that once seized carts "for the walls." The whispered judgments in Synod Hall are no longer just legal proceedings, but something far more sinister.

The entity feeds on the accumulated fears and secrets of centuries of scholars, students, and seekers who have passed through these halls, making every corridor a potential trap and every book a source of both wisdom and terror. The very air seems to thicken with unspoken knowledge, and shadows move with purpose that defies natural law.

\subsection{Key Characteristics}

\begin{itemize}
\item \textbf{Architectural Horror:} The very building defies logic, with corridors that shift and rooms that are larger within than without. Staircases spiral into themselves, leading not upward but inward, toward some geometrical heart of madness.
\item \textbf{Isolation:} Characters are cut off from help within the University complex, trapped in a maze of stone and shadow where every door leads to another mystery and every window shows only the endless void of night. The outside world becomes a distant memory.
\item \textbf{Psychological Warfare:} The enemy cannot be easily understood or fought, operating on mental rather than physical levels. Reality itself becomes a weapon, bending and breaking around the edges of perception.
\item \textbf{Escalating Threat:} Fear builds throughout the campaign as the entity grows stronger, its whispers becoming roars, its shadows becoming substance, until the boundary between mind and world dissolves entirely.
\item \textbf{Moral Complexity:} Survival may require compromising principles and making difficult ethical choices. The price of knowledge is paid in sanity, and the cost of truth may be the soul itself.
\end{itemize}

\subsection{Navigating the University}

Navigation in Thepyrgos is not merely about finding one's way through physical spaces, but understanding the psychological landscape of fear and knowledge. The Deck System becomes crucial here, as each draw represents not just a random encounter but a meaningful narrative element that shapes the growing horror. The university breathes, and with each exhalation, new corridors manifest while others collapse into memory.

\section{Key NPCs and Entities}

\subsection{Aqyl, Son of Aqyl}

The Enigmatic Scholar who serves as both guide and potential threat in the depths of academic pursuit.

\begin{monsterentry}{Aqyl, Son of Aqyl}
\textbf{Threat Level:} Major \\
\textbf{Harm:} Varies \\
\textbf{SB:} 2–3 SB per interaction \\
\textbf{Description:} Aqyl appears as a distinguished scholar with deep knowledge of the University's hidden passages and ancient texts. His eyes reflect unusual colors in certain lighting, shifting from scholarly brown to depths that seem to contain entire galaxies. He speaks in whispers that carry more than words — each syllable seems to echo from vast distances, as if his voice originates from the spaces between stars. \\
\textbf{Motivations:} Maintain the University's secrets while protecting those he cares about \\
\textbf{Abilities:}
\begin{itemize}
    \item Extensive Knowledge: Deep understanding of hidden passages and ancient texts
    \item Whispered Communication: Speaks in whispers that carry supernatural influence
    \item Oath-Bound Wisdom: Cannot easily share certain forbidden knowledge
    \item Partial Influence Resistance: Shows early signs of entity's touch
\end{itemize}
\textbf{Weaknesses:}
\begin{itemize}
    \item Bound by oaths and knowledge he cannot unlearn
    \item Vulnerable to direct confrontation about his true nature
    \item Becomes confused by absolute statements about reality
\end{itemize}
\textbf{Sample Encounter:} Aqyl offers guidance through the ancient stacks but his whispered advice leads to places that test the PCs' understanding of their own sanity. The GM can spend their SB to:
\begin{itemize}
    \item 1 SB: His guidance reveals useful but disturbing information
    \item 2 SB: He begins to show physical signs of corruption — his shadow moves independently
    \item 3 SB: His whispers contain hidden commands that bypass conscious resistance
    \item 4+ SB: He becomes a conduit for the entity's influence, his body a vessel for cosmic horror
\end{itemize}
\end{monsterentry}

\subsection{Palikar Captain Thorne}

The Reluctant Guardian whose duty conflicts with his growing fear of what lurks within the University walls.

\begin{monsterentry}{Palikar Captain Thorne}
\textbf{Threat Level:} Major \\
\textbf{Harm:} \textgreater \\
\textbf{SB:} 2–3 SB per scene \\
\textbf{Description:} Captain Thorne appears as a weathered military officer with knowledge of tower defenses and patrol routes. His uniform, once crisp and authoritative, now bears stains that seem to shift and writhe in peripheral vision. He has seen colleagues disappear into the stacks, never to return, and his letter-shyness masks deep paranoia that has begun to manifest physically — his hands shake when he speaks of the deeper levels. \\
\textbf{Motivations:} Protect the University from external threats while dealing with internal corruption \\
\textbf{Abilities:}
\begin{itemize}
    \item Tactical Knowledge: Understanding of tower defenses and patrol routes
    \item Military Training: Combat expertise and tactical awareness
    \item Paranoid Awareness: Heightened sense of supernatural danger
    \item Authority Command: Can rally remaining guards and scholars
\end{itemize}
\textbf{Weaknesses:}
\begin{itemize}
    \item Letter-shy and reluctant to share information
    \item Becoming increasingly erratic under supernatural pressure
    \item Vulnerable to psychological manipulation
\end{itemize}
\textbf{Sample Encounter:} Captain Thorne approaches the PCs offering protection but his military precision begins to crack under the entity's influence. The GM can spend their SB to:
\begin{itemize}
    \item 1 SB: He reveals knowledge of disappeared colleagues, his voice dropping to a terrified whisper
    \item 2 SB: His paranoia causes him to suspect the PCs, drawing his weapon with trembling hands
    \item 3 SB: He becomes partially corrupted and hunts the PCs, his eyes reflecting the entity's light
    \item 4+ SB: He sacrifices himself to buy time for the PCs, his final words a warning whispered from beyond death
\end{itemize}
\end{monsterentry}

\section{Horror Elements and Dread Mechanics}

\subsection{Dread as Resource Management}

The University of Thepyrgos is inherently hostile to the human psyche. The isolation, the unknown threats, and the constant psychological warfare all contribute to a slow erosion of sanity that must be carefully managed. The very walls seem to press inward, and the silence between heartbeats grows longer with each passing moment.

\begin{mechanic}{Dread Clock}
Instead of traditional sanity points, players must spend Boons to prevent the Dread Clock from advancing. This creates a resource management challenge where players must choose between immediate tactical advantages and long-term psychological stability. Each tick of the clock brings them closer to a threshold where reality itself becomes unreliable.
\end{mechanic}

\subsection{The Dread Clock}

\begin{center}
\begin{tabular}{|m{4cm}|m{8cm}|}
\hline
\rowcolor{tableheader}
\textbf{Dread Clock} & \textbf{Psychological deterioration and mounting horror} \\
\hline
Segments & \textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet 0/10 \\
\hline
\end{tabular}
\end{center}

\subsubsection{Advancement Triggers}

\begin{itemize}
\item Discovering scholars' fate: +1 segment (prevent with 1 Boon) — their final notes speak of things that should not be known
\item Hearing whispers in the dark: +1 segment (prevent with 1 Boon) — voices that seem to originate from within your own skull
\item Seeing shadows move unnaturally: +1 segment (prevent with 1 Boon) — figures that cast no light but absorb all shadow
\item Finding evidence of entity's influence: +2 segments (prevent with 2 Boons) — symbols that hurt to look upon directly
\item Companion shows signs of corruption: +2 segments (prevent with 2 Boons) — their eyes have begun to reflect impossible depths
\item Direct psychic attack from entity: +3 segments (prevent with 3 Boons) — a presence that knows your deepest fears and speaks them aloud
\end{itemize}

\subsubsection{Psychological Effects by Dread Level}

\begin{itemize}
\item \textbf{0-2 Segments - Unease:} Characters feel on edge, -1 die to social rolls involving trust, minor sleep disturbances. The walls seem to breathe, and footsteps echo from empty corridors.
\item \textbf{3-4 Segments - Fear:} -1 die to all rolls when alone, startle easily, avoid dark spaces, nightmares. Shadows have begun to gather at the edges of vision, watching with intent.
\item \textbf{5-6 Segments - Terror:} -2 dice to rolls involving courage or rational thought, may flee from non-threatening situations. The air grows thick with unspoken words, and reality begins to fray at the edges.
\item \textbf{7-8 Segments - Madness:} -2 dice to most rolls, may act against own interests, hallucinations, unreliable perception. Time flows backward in certain corridors, and books rewrite themselves when not observed.
\item \textbf{9-10 Segments - Broken:} Out of control, may become hostile to allies, permanent psychological damage, requires major intervention to recover. The character has become a conduit for cosmic knowledge they cannot comprehend.
\end{itemize}

\section{Key Locations in Thepyrgos}

\subsection{Synod Hall}

The heart of the University's legal and mystical authority, where judgment is audible at a whisper. The hall stretches impossibly far in all directions, its ceiling lost in shadow. Massive tomes float in the air, their pages turning of their own accord, recording judgments that affect not just the living, but the very fabric of reality itself. The whispers here are not human voices, but the susurrus of cosmic law being rewritten.

\subsection{The Forbidden Stacks}

Ancient library levels where forbidden knowledge is kept, and where the entity's influence is strongest. The books here are not bound by conventional physics — they shift and rearrange themselves, their spines displaying titles in languages that predate human speech. The air is thick with the scent of old parchment and something else, something that smells of ozone and the void between stars. Reading here is dangerous, as knowledge has weight and some truths are too heavy for mortal minds to bear.

\subsection{The North Tower}

Abandoned tower where the first signs of corruption appeared, serving as a focal point for the entity's influence. The tower leans at an impossible angle, defying gravity with casual arrogance. Windows show not the outside world, but glimpses of other realities — some beautiful beyond description, others so terrible that looking upon them causes immediate psychological trauma. The stairs spiral upward into darkness, and those who climb them often find themselves returning to the same landing, no matter which direction they travel.

\subsection{The Palikar Barracks}

Guard quarters where the University's protectors have become its prisoners, filled with evidence of disappearances. The rooms are perfectly maintained, as if the occupants stepped out for a moment and never returned. Personal effects remain — half-drunk tea cups, books left open at specific pages, weapons still in their racks. But the mirrors show only empty space, and the shadows in the corners move when no one is looking.

\section{Custom Horror Mechanics}

\subsection{The Whispering Mechanic}

When in the ancient towers or stacks of the University, PCs must make Wits + Lore rolls (DV 3) to resist hearing the entity's whispers. Each failure:
\begin{itemize}
\item Generates 1 SB that the GM can spend for psychological effects
\item Advances Dread Clock by 1 segment (prevent with 1 Boon)
\item May reveal useful but disturbing information
\end{itemize}

\textbf{Whisper Examples:}
\begin{itemize}
\item "The books remember your name, and they are hungry..."
\item "Knowledge has a price, and it is always more than you can pay..."
\item "The Matriarch waits for you in Synod Hall, where all judgments are final..."
\item "Your companion's thoughts are not their own — count the shadows that follow them..."
\item "The geometry of this place is wrong. You are not where you think you are..."
\end{itemize}

\subsection{Sacred Geometry Perception}

When PCs observe the ancient architecture of the University, particularly in Synod Hall or the older towers, they must make Wits + Investigation rolls (DV 4) to avoid comprehension effects. Each failure:
\begin{itemize}
\item Generates 2 SB that the GM can spend for reality distortions
\item Advances Dread Clock by 2 segments (prevent with 2 Boons)
\item May grant forbidden knowledge at great psychological cost
\end{itemize}

\textbf{Geometry Manifestations:}
\begin{itemize}
\item Corridors that should be straight but bend impossibly, leading not to other rooms but to other dimensions
\item Rooms that are larger on the inside than the outside, their walls stretching into infinite space
\item Symbols that shift when not directly observed, rearranging themselves into new patterns of meaning
\item Stairs that lead to different floors depending on the direction of approach, creating temporal loops
\item Doors that open onto themselves, creating portals to the spaces between thoughts
\end{itemize}

\subsection{Knowledge Corruption}

PCs who reach 7+ Dread segments begin to show physical signs of the entity's influence:
\begin{itemize}
\item Eyes that reflect unusual colors in darkness — deep purples, impossible golds, the black of space between stars
\item Speaking in whispers without realizing it, their voice carrying the weight of cosmic knowledge
\item Attraction to dark, enclosed spaces like the ancient stacks, as if drawn by some gravitational pull
\item May be able to communicate with the entity through forbidden knowledge, their words becoming prophecy
\end{itemize}

\textbf{Corruption Effects:}
\begin{itemize}
\item +1 die to Lore rolls involving forbidden knowledge, but each use advances the Dread Clock by 1
\item -1 die to social rolls due to unsettling presence — others instinctively avoid the character
\item Can perceive multiple timeline branches (generates 2 SB per scene) — the weight of infinite possibilities
\item Permanent reality distortion (narrative consequence) — the character exists partially outside normal reality
\end{itemize}

\subsection{The Collective Dread}

The party's collective Dread affects their perception of the University:
\begin{itemize}
\item Average Dread level determines reality stability — higher levels make the impossible more probable
\item High average = shared hallucinations, impossible events become real, the university reshapes itself around their fears
\item Low average = grounding effect, some resistance to cosmic influence, reality maintains a tenuous hold
\end{itemize}

\textbf{Collective Effects:}
\begin{itemize}
\item Shared visions of the entity's true form — a geometry of madness that breaks the mind
\item Impossible architectural changes that affect all PCs — corridors that didn't exist before suddenly appear
\item Collective memory gaps about recent events — time itself becomes unreliable
\item Enhanced group paranoia and infighting — trust becomes a luxury they cannot afford
\end{itemize}

\section{Campaign Clocks}

\subsection{Entity's Awakening Clock (12 segments)}

\begin{center}
\begin{tabular}{|m{4cm}|m{8cm}|}
\hline
\rowcolor{tableheader}
\textbf{Entity's Awakening Clock} & \textbf{Progress toward the collective consciousness of forbidden knowledge being fully disturbed} \\
\hline
Segments & \textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet 0/12 \\
\hline
\end{tabular}
\end{center}

\textbf{Advancement Triggers:}
\begin{itemize}
\item Forbidden texts read: +1 segment per major tome — each book is a doorway to madness
\item Ancient secrets uncovered: +2 segments — knowledge that predates human civilization
\item Synod Hall judgment heard: +1 segment — cosmic law being rewritten in real-time
\item Knowledge used for dark purposes: +2 segments — the entity grows stronger with each application
\item PCs delve deeper into forbidden stacks: +1 segment per session — descending into the heart of madness
\item PCs interfere with awakening: +2 segments — resistance feeds the entity's hunger
\item Ancient binding ritual discovered: -1 segment — hope flickers in the darkness
\end{itemize}

\subsection{Town Collapse Clock (8 segments)}

\begin{center}
\begin{tabular}{|m{4cm}|m{8cm}|}
\hline
\rowcolor{tableheader}
\textbf{Town Collapse Clock} & \textbf{How quickly the University community breaks down under supernatural pressure} \\
\hline
Segments & \textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet 0/8 \\
\hline
\end{tabular}
\end{center}

\textbf{Advancement Triggers:}
\begin{itemize}
\item Dread Clock advances: +1 segment — madness is contagious
\item Townspeople disappear or go mad: +1 segment each — the population dwindles
\item PCs fail to provide leadership: +1 segment — chaos feeds on uncertainty
\item Supernatural events witnessed by townsfolk: +2 segments — reality itself becomes unreliable
\item Essential services fail: +1 segment — the infrastructure of civilization crumbles
\end{itemize}

\subsection{Whispering Stacks Clock (8 segments)}

\begin{center}
\begin{tabular}{|m{4cm}|m{8cm}|}
\hline
\rowcolor{tableheader}
\textbf{Whispering Stacks Clock} & \textbf{Progress toward the entity's full manifestation through accumulated knowledge} \\
\hline
Segments & \textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet\textbullet 0/8 \\
\hline
\end{tabular}
\end{center}

\textbf{Advancement Triggers:}
\begin{itemize}
\item Forbidden texts read: +1 segment per major tome — each book is a sacrifice to the entity
\item Ancient secrets uncovered: +2 segments — knowledge that should remain buried
\item Synod Hall judgment heard: +1 segment — legal decisions with cosmic consequences
\item Knowledge used for dark purposes: +2 segments — the entity grows stronger with each revelation
\item PCs delve deeper into forbidden stacks: +1 segment per session — descending into the abyss
\end{itemize}

\section{Sample Sessions}

\subsection{Session 1: Arrival at Thepyrgos}

\textbf{Opening Scene:} The PCs arrive to find the University eerily quiet with most towers abandoned. The silence is not peaceful, but oppressive — the kind of quiet that comes when all sound is absorbed by hungry shadows.

\textbf{Key Encounters:}
\begin{enumerate}
\item Investigation of the abandoned North Tower (Wits + Investigation) — the tower leans at an impossible angle, defying all laws of physics
\item Conversation with the increasingly paranoid porter (Presence + Sway) — his eyes reflect depths that should not exist
\item First encounter with whispers in the dark stacks (Wits + Perception, DV 3) — voices that seem to originate from within your own mind
\item Discovery of strange symbols carved near the Synod Hall entrance — symbols that hurt to look upon directly
\end{enumerate}

\textbf{Dread Clock Advancement:}
\begin{itemize}
\item First whisper encounter: +1 segment (prevent with 1 Boon) — the voices know your name
\item Seeing abandoned, obviously terrified scholar: +2 segments (prevent with 2 Boons) — their final notes speak of things that should not be
\item Discovering symbols that shouldn't exist: +1 segment (prevent with 1 Boon) — geometry that breaks the mind
\end{itemize}

\textbf{Campaign Clock Advancement:}
\begin{itemize}
\item Entity's Awakening: +1 (general unease in the University) — the entity stirs in its ancient slumber
\item Town Collapse: +1 (porter's nervousness) — madness spreads like a contagion
\end{itemize}

\subsection{Session 2: Descent into Darkness}

\textbf{Key Encounters:}
\begin{enumerate}
\item Exploration of the ancient stacks beneath the University (Wits + Survival) — corridors that lead not to other rooms but to other dimensions
\item Encounter with a corrupted scholar who speaks in whispers (combat + social) — their words carry the weight of cosmic knowledge
\item Discovery of the ritual chamber deep in the forbidden stacks (Wits + Lore) — a space where reality is negotiable
\item First direct contact with entity's influence through whispered knowledge (Spirit + Resolve, DV 5) — the presence that knows your deepest fears
\end{enumerate}

\textbf{Dread Clock Advancement:}
\begin{itemize}
\item Seeing first corrupted scholar: +2 segments (prevent with 2 Boons) — their eyes reflect impossible depths
\item Being touched by entity's influence: +3 segments (prevent with 3 Boons) — knowledge that rewrites the soul
\item Discovering extent of corruption: +2 segments (prevent with 2 Boons) — the university is already lost
\end{itemize}

\textbf{Campaign Clock Advancement:}
\begin{itemize}
\item Entity's Awakening: +2 (seals disturbed) — ancient protections begin to fail
\item Town Collapse: +1 (more scholars disappear) — the population dwindles
\end{itemize}

\subsection{Session 3: The Truth Revealed}

\textbf{Key Encounters:}
\begin{enumerate}
\item Research in Aqyl's notes (Wits + Lore) — knowledge that predates human civilization
\item Confrontation with Palikar Captain Thorne as he becomes erratic (Presence + Command) — his final words are a warning from beyond death
\item Discovery of ancient warding techniques in the vaults (Wits + Arcana) — protections that may yet hold
\item Choice: Attempt to reinforce seals or flee while there's still time — every moment of hesitation feeds the entity
\end{enumerate}

\textbf{Dread Clock Advancement:}
\begin{itemize}
\item Learning the true nature of the entity: +3 segments (prevent with 3 Boons) — a geometry of madness that breaks the mind
\item Witnessing Captain Thorne's breakdown: +2 segments (prevent with 2 Boons) — the last line of defense crumbles
\item Realizing the scope of the threat: +2 segments (prevent with 2 Boons) — the entity is older than human memory
\end{itemize}

\textbf{Campaign Clock Advancement:}
\begin{itemize}
\item Entity's Awakening: +3 (major seal damaged) — the entity stirs with cosmic hunger
\item Town Collapse: +2 (Captain's authority breaks down) — chaos reigns in the streets
\end{itemize}

\section{Resolution Paths}

\subsection{The Sacrifice}

Permanently seal the entity using ancient techniques, but it requires one PC to remain behind as a living anchor. The character becomes part of the university itself, their consciousness spread throughout the stone and shadow, forever whispering warnings to future seekers of knowledge.

\begin{itemize}
\item Success means the entity is contained, but at great personal cost — the character exists in a state between life and death
\item Award 15–18 XP
\item The sacrificed PC becomes a guardian spirit, occasionally communicating through whispers that carry fragments of cosmic truth
\item The University remains but is forever changed — some areas are permanently sealed, others become sanctuaries of terrible knowledge
\end{itemize}

\subsection{The Escape}

Flee with evidence of the threat, warning other settlements. The entity remains but is contained for now, though its influence spreads slowly to neighboring regions like ink through water. The PCs become hunted by those who want to suppress the truth, and haunted by memories of what they witnessed in the depths.

\begin{itemize}
\item Award 10–12 XP
\item Create ongoing campaign thread — the entity's influence spreads like a contagion
\item The University becomes a quarantined zone, sealed by those who fear the truth
\item The entity's influence spreads slowly to neighboring regions, corrupting knowledge wherever it takes root
\item PCs become hunted by those who want to suppress the truth, and haunted by what they cannot forget
\end{itemize}

\subsection{The Corruption}

Allow the entity to partially manifest, gaining its power but becoming its servants. The PCs become conduits for cosmic knowledge they cannot fully comprehend, their humanity slowly eroded by the weight of infinite possibilities. They gain power, but at the cost of their souls.

\begin{itemize}
\item Transform PCs into agents of horror — their words carry the weight of prophecy
\item Award 8–10 XP but fundamentally change character nature — they exist partially outside normal reality
\item PCs gain supernatural abilities but lose humanity — their eyes reflect impossible depths
\item They become extensions of the entity's will — knowledge flows through them like blood through veins
\item The University becomes a hub for the entity's expansion — a nexus of cosmic influence
\end{itemize}

\subsection{The Investigation}

Fully understand the entity and find a way to banish it without sacrifice. This path requires significant research and resources, and success comes at the cost of confronting truths that may be too terrible to bear. The PCs must gather knowledge from multiple forbidden texts, each one a step closer to madness.

\begin{itemize}
\item Requires significant research and resources — knowledge that predates human civilization
\item Award 18–20 XP if successful, but very difficult — the price of truth is always high
\item Must gather knowledge from multiple forbidden texts — each book is a doorway to madness
\item Requires cooperation with corrupted NPCs — trust becomes a luxury they cannot afford
\item Success permanently seals the entity but weakens the fabric of reality in the area — some truths are too heavy for the world to bear
\end{itemize}

\subsection{The Bargain}

Negotiate with the entity to limit its influence in exchange for specific concessions. The entity agrees to limit its feeding in exchange for periodic offerings — knowledge, souls, or fragments of reality itself. The PCs become mediators between the entity and the living, walking a razor's edge between cosmic horror and human sanity.

\begin{itemize}
\item Award 12–15 XP with ongoing supernatural responsibilities — the burden of knowledge never ends
\item The entity agrees to limit its feeding in exchange for periodic offerings — knowledge, souls, or reality itself
\item PCs become mediators between the entity and the living — walking a razor's edge between horror and sanity
\item The University becomes a neutral ground for otherworldly negotiations — a place where cosmic law is rewritten daily
\item Creates potential for future conflicts when the bargain is tested — every concession feeds the entity's hunger
\end{itemize}

\section{Using the Deck System}

\subsection{Drawing from Thepyrgos Deck}

The Deck System provides rich narrative content for encounters within the University. Each draw should feel meaningful and contribute to the overall atmosphere of paranoia and unknown threats. The cards are not random events, but manifestations of the entity's will, shaped by the collective fears and desires of those who have come before.

\subsubsection{Hearts - Emotional/Social Fallout}

\begin{itemize}
\item Paranoia spreads among remaining scholars — trust becomes a luxury they cannot afford
\item PCs turn on each other due to whispered suggestions — the entity speaks through their fears
\item A trusted NPC reveals they've been compromised — their eyes reflect impossible depths
\item Romantic subplot becomes complicated by supernatural influence — love and madness are indistinguishable
\item Family connections are used against PCs by the entity — the past becomes a weapon
\end{itemize}

\subsubsection{Spades - Harm/Escalation}

\begin{itemize}
\item Physical manifestations of knowledge corruption — reality itself becomes a weapon
\item Architecture shifts to trap or harm PCs — the university reshapes itself around their fears
\item Corrupted scholars become hostile — their words carry the weight of cosmic knowledge
\item Ancient defenses activate against intruders — protections that predate human civilization
\item Reality distortions cause physical injury — the laws of physics are negotiable
\end{itemize}

\subsubsection{Clubs - Resource Depletion}

\begin{itemize}
\item Essential supplies become contaminated — food that tastes of ash and knowledge
\item Communication with outside world is cut off — the university exists in a bubble of isolation
\item Magical or technological aids malfunction — reality itself resists their intrusion
\item Time pressure as the entity's awakening accelerates — every moment of hesitation feeds its hunger
\item Allies become unavailable due to corruption or disappearance — trust no one
\end{itemize}

\subsubsection{Diamonds - Magical/Spiritual Disturbance}

\begin{itemize}
\item Forbidden knowledge reveals itself unexpectedly — truths that break the mind
\item The entity manifests partially in the physical world — geometry that should not exist
\item Ancient wards begin to fail — protections that have held for millennia crumble
\item Reality itself becomes unstable — the laws of physics bend and break
\item Supernatural entities from other dimensions take notice — the university is a beacon of cosmic significance
\end{itemize}

\section{GM Tips and Advice}

\subsection{Atmosphere and Tension}

Building Dread:
\begin{itemize}
\item Use lighting, sound, and physical environment to create unease — shadows that move independently, whispers that seem to originate from within the walls
\item Describe sensations and feelings, not just visual details — the air grows thick with unspoken words, the silence between heartbeats grows longer
\item Let silences and pauses carry weight — sometimes what is not said is more terrifying than what is
\item Make the familiar seem alien and threatening — the university was built by minds that comprehended geometries beyond human understanding
\end{itemize}

Pacing the Horror:
\begin{itemize}
\item Start subtle and build gradually — the first whispers are almost comforting, like old friends speaking in the dark
\item Vary the intensity - allow moments of false security — the university can be beautiful in its terrible way
\item Use foreshadowing and ominous signs — symbols that shift when not directly observed, corridors that lead to places that should not exist
\item Save the biggest revelations for climactic moments — the entity's true form is a geometry of madness that breaks the mind
\end{itemize}

\subsection{Managing Fear and Sanity}

Fear as a Resource:
\begin{itemize}
\item Fear should be a narrative driver, not a punishment — it shapes the story, not just the characters
\item Let players feel powerful even when afraid — knowledge is power, even when it comes at a terrible cost
\item Provide opportunities to confront and overcome fears — sometimes the only way forward is through the heart of madness
\item Balance terror with moments of triumph — even in the darkest places, hope can flicker like a candle in the void
\end{itemize}

Sanity Management:
\begin{itemize}
\item Make sanity loss feel meaningful and personal — each loss is a piece of the character's humanity
\item Let it change how characters perceive and interact with the world — reality becomes negotiable
\item Provide ways to recover or adapt to mental trauma — some wounds heal, others become scars that define the character
\item Avoid making characters useless when sanity is low — madness can be a form of knowledge
\end{itemize}

\subsection{Narrative Techniques}

Player Agency:
\begin{itemize}
\item Give players meaningful choices, even when options seem limited — every decision feeds the entity, but some feed it more than others
\item Let their decisions have real consequences — the university remembers everything
\item Provide multiple approaches to problems — sometimes the only way forward is through knowledge, sometimes through violence
\item Respect their courage to face the horror head-on — bravery in the face of cosmic insignificance is its own form of heroism
\end{itemize}

Unreliable Information:
\begin{itemize}
\item Not all sources are trustworthy — the entity speaks through every voice
\item Clues may be misleading or incomplete — knowledge is a weapon that cuts both ways
\item NPCs might be lying, mistaken, or hiding crucial information — trust is a luxury they cannot afford
\item Let players discover the truth through investigation and deduction — the university rewards those who seek understanding
\end{itemize}

\section{Appendix: Additional Resources}

\subsection{Character Options}

\textbf{Recommended Backgrounds:}
\begin{itemize}
\item Scholar of Fractured Truths (Wizard archetype) — one who has glimpsed forbidden knowledge and lived to regret it
\item The Chronicler of Consequences (Bard archetype) — one who records the price of every choice
\item The Caretaker of Cycles (Druid archetype) — one who understands that all things must end
\item The Guild-Approved Shadow (Rogue archetype) — one who moves through darkness without being consumed
\item The Border-Warden (Ranger archetype) — one who patrols the spaces between worlds
\end{itemize}

\textbf{Useful Skills:}
\begin{itemize}
\item Lore (Essential for understanding the entity) — knowledge that predates human civilization
\item Investigation (Key for discovering clues) — the ability to see patterns in chaos
\item Arcana (For dealing with supernatural elements) — magic that bends reality itself
\item Insight (To detect corruption in others) — the ability to see the truth behind the mask
\item Survival (For navigating the dangerous stacks) — the university is a maze with no exit
\item Diplomacy (For dealing with NPCs) — words can be weapons more terrible than any blade
\item Stealth (For avoiding corrupted entities) — sometimes the only way to survive is to remain unseen
\end{itemize}

\textbf{Suggested Talents:}
\begin{itemize}
\item Lorekeeper (Recall obscure history or magic) — memory is a form of immortality
\item Backlash Soothing (Reduce magical Backlash) — some wounds heal, others become scars
\item Silver Tongue (Persuade through speech) — words can reshape reality itself
\item Battle Instincts (Re-roll failed defense rolls) — survival is its own form of wisdom
\item Iron Stomach (Resist mundane poisons) — the body must be strong to house a mind that has seen too much
\item Exceptional Coordination (Follower provides +4 assist dice) — trust is a luxury, but sometimes necessary
\end{itemize}

\subsection{Protective Items}

\begin{itemize}
\item \textbf{Blessed Warding Stone:} Provides +2 dice to resist supernatural influence and reduces Dread Clock advancement by 1 (minimum 1). Crumbles to dust if Dread Clock fills completely. The stone whispers warnings in a language that predates human speech.
\item \textbf{Sanctified Salt:} Can be used to create protective circles or barriers. One use per session to prevent Dread Clock advancement or gain start Controlled vs. supernatural threats. The salt burns with cold fire when the entity is near.
\item \textbf{Prayer Book of the Steadfast:} Once per session, allows reroll of failed Spirit + Resolve roll and prevents 1 segment of Dread Clock advancement. The pages turn of their own accord, revealing passages that speak to the character's deepest fears.
\end{itemize}

\subsection{Cursed Artifacts}

\begin{itemize}
\item \textbf{Whispering Tome:} A book that hums with voices inaudible to most. Reader gains +1 die to Lore rolls but must make Spirit + Resolve (DV 3) each session or advance Dread Clock by 1. The voices speak of things that should not be known, and their words carry the weight of cosmic truth.
\item \textbf{Mirror of the Forgotten:} Shows glimpses of possible futures and past events. Provides valuable insights (Wits + Lore, DV 3) but each use advances Dread Clock by 1 and generates 1 SB. The mirror shows not reflection, but possibility — and some possibilities are too terrible to contemplate.
\item \textbf{Shroud of Whispers:} A burial cloth that grants immunity to fear effects but causes the wearer to hear the entity's whispers constantly. Must make Wits + Perception (DV 4) or advance Dread Clock by 2 each day worn. The shroud is woven from the dreams of the dead, and its wearer becomes a conduit for cosmic knowledge.
\end{itemize}

\section{Conclusion}

"Whispers in the Stacks" is designed to challenge your players not just physically, but psychologically and morally. The horror lies not in gore or jump scares, but in the slow erosion of certainty and the terrifying realization that knowledge itself can be a curse. The university is a place where reality is negotiable and truth comes at a price that can never be fully paid.

Remember that the best horror comes from what is left unsaid and unseen. Let the players' imaginations fill in the gaps, and trust in the power of implication over explicit description. The fear of what might be lurking in the shadows is often more effective than showing every monster in detail. The entity is not just a threat — it is a presence that reshapes reality itself, making the impossible seem inevitable.

Most importantly, maintain communication with your players about comfort levels and boundaries. Horror can be deeply personal, and what one player finds thrilling, another might find genuinely distressing. A good horror game is one where everyone at the table is having fun and feeling appropriately challenged, not traumatized.

The mechanics provided here are tools to enhance the horror experience, not replace good storytelling and atmosphere. Use them to support your narrative goals and create memorable, chilling experiences for your players.

As the GM, you hold the quill that writes the legend, but it is the players who create the story. Guide them through the stacks of Thepyrgos, challenge their perceptions, and let them emerge changed by their journey - for better or for worse.

\begin{center}
\textbf{In the spaces between knowledge, in the whispers of the stacks, and in the judgments of Synod Hall, the truth waits patiently. Will you listen? Will you pay the price? Will you survive to tell the tale?}
\end{center}

The University of Thepyrgos thanks you for your service. Please return all forbidden texts to the proper authorities. Knowledge is a responsibility, not a right.

May your dice roll true, your choices be meaningful, and your players emerge from the stacks forever changed.

\newpage

\section*{Quick Reference Cards}

\subsection*{Dread Clock Management}

\begin{tabular}{|p{5cm}|p{7cm}|
\hline
\rowcolor{tableheader}
\textbf{Trigger} & \textbf{Segments} \\
\hline
Discovering scholars' fate & +1 (prevent with 1 Boon) \\
Hearing whispers in the dark & +1 (prevent with 1 Boon) \\
Seeing shadows move unnaturally & +1 (prevent with 1 Boon) \\
Finding evidence of entity's influence & +2 (prevent with 2 Boons) \\
Companion shows signs of corruption & +2 (prevent with 2 Boons) \\
Direct psychic attack from entity & +3 (prevent with 3 Boons) \\
\hline
\end{tabular}

\subsection*{Psychological Effects}

\begin{tabular}{|p{2cm}|p{10cm}|}
\hline
\rowcolor{tableheader}
\textbf{Segments} & \textbf{Effects} \\
\hline
0-2 & Unease: -1 die to social rolls involving trust, minor sleep disturbances \\
3-4 & Fear: -1 die to all rolls when alone, startle easily, avoid dark spaces \\
5-6 & Terror: -2 dice to rolls involving courage or rational thought, may flee \\
7-8 & Madness: -2 dice to most rolls, may act against own interests, hallucinations \\
9-10 & Broken: Out of control, may become hostile, permanent psychological damage \\
\hline
\end{tabular}

\subsection*{Campaign Clocks Quick Reference}

\textbf{Entity's Awakening Clock (12 segments):}
\begin{itemize}
\item Measures progress toward collective consciousness disturbance
\item Triggers: Forbidden texts, ancient secrets, Synod judgments
\end{itemize}

\textbf{Town Collapse Clock (8 segments):}
\begin{itemize}
\item Tracks University community breakdown
\item Triggers: Dread advancement, disappearances, service failures
\end{itemize}

\textbf{Whispering Stacks Clock (8 segments):}
\begin{itemize}
\item Measures entity's manifestation through accumulated knowledge
\item Triggers: Forbidden texts, ancient secrets, Synod judgments
\end{itemize}

\subsection*{Custom Mechanics Summary}

\textbf{Whispering Mechanic:}
\begin{itemize}
\item Wits + Lore (DV 3) to resist entity whispers
\item Failure: 1 SB, +1 Dread segment, disturbing information
\end{itemize}

\textbf{Sacred Geometry Perception:}
\begin{itemize}
\item Wits + Investigation (DV 4) to avoid comprehension effects
\item Failure: 2 SB, +2 Dread segments, forbidden knowledge
\end{itemize}

\textbf{Knowledge Corruption (7+ Dread):}
\begin{itemize}
\item Physical signs: unusual eye colors, whispering, attraction to dark spaces
\item Effects: +1 Lore die, -1 social die, timeline perception, reality distortion
\end{itemize}

\subsection*{Deck-Based Complications}

\textbf{Hearts (Social/Emotional):}
\begin{itemize}
\item Paranoia spreads, PCs turn on each other, NPC compromise
\end{itemize}

\textbf{Spades (Harm/Escalation):}
\begin{itemize}
\item Physical corruption, architectural traps, hostile scholars
\end{itemize}

\textbf{Clubs (Resources):}
\begin{itemize}
\item Contaminated supplies, communication cut, aid malfunction
\end{itemize}

\textbf{Diamonds (Supernatural):}
\begin{itemize}
\item Forbidden knowledge, partial manifestation, ward failure
\end{itemize}

\subsection*{Resolution Paths Summary}

\begin{tabular}{|p{4cm}|p{8cm}|}
\hline
\rowcolor{tableheader}
\textbf{Path} & \textbf{Outcome \& XP} \\
\hline
The Sacrifice & Seal entity with PC as anchor. Award 15–18 XP. \\
The Escape & Flee with evidence. Award 10–12 XP. \\
The Corruption & Become entity's servants. Award 8–10 XP. \\
The Investigation & Banish entity through research. Award 18–20 XP. \\
The Bargain & Negotiate entity limitations. Award 12–15 XP. \\
\hline
\end{tabular}

\end{document}

