% --- Fate's Edge SRD --- Section 1: Core Principles (Revised) ---
% Include this file from your main .tex with: \input{01-core-principles.tex}

\section{Core Principles}

% --- Safety Tools: Lines, Veils & X-Card ---
% Place this after the title page or as Section 0 of your SRD.
\begin{tcolorbox}[
  breakable,
  colback=red!5!white,
  colframe=red!75!black,
  title=\textbf{Safety Tools: Lines, Veils \& X-Card},
  fonttitle=\bfseries\large,
  sharp corners,
  drop shadow,
  before skip=10pt,
  after skip=10pt
]

\subsection*{X-Card}
Any player (including the GM) can draw an "X" on a card (or simply say "X") when a moment makes them uncomfortable. 
When the X‑Card is shown, the described content is \textbf{removed from the fiction} -- no questions, no explanations, no judgments. 
You do not need to justify using the X‑Card. The scene is retconned or redirected immediately.

\subsection*{Lines}
A \textbf{Line} is a hard boundary: content that will not appear in the game at all. 
Examples: "No sexual violence," "No harm to children," "No body horror." 
Before the campaign or a session, players name their Lines. The GM and group honor them completely.

\subsection*{Veils}
A \textbf{Veil} is a soft boundary: content that can happen "off screen" or fade to black. 
The event is acknowledged but not described in detail. 
Example: "A romantic scene can happen, but fade to black before it becomes explicit." 
Veils allow the story to acknowledge mature themes without forcing players to roleplay through them.

\subsection*{How to Use These Tools at Your Table}
\begin{itemize}
  \item Establish Lines and Veils during \textbf{Session Zero} (or before your first game). Write them down where everyone can see them.
  \item Keep an X‑Card visible on the table (a printed card, a drawn X, or a digital symbol).
  \item Remind players they can use the X‑Card at any time, even mid‑sentence.
  \item When the X‑Card is used, \textbf{pause}, rewind if needed, and move on. Do not interrogate the player who used it.
  \item The GM is also a player. The GM may use the X‑Card for their own comfort.
  \item Respect the tools. If someone calls an X‑Card, trust that they had a good reason and support them.
\end{itemize}

\subsection*{Why These Tools Matter}
These tools are not about censorship -- they are about \textbf{consent}. 
They allow the group to tell darker, more intense stories together because everyone knows they can tap out at any time. 
A game that feels safe is a game that can go to the edge.

\smallskip
\noindent\textit{For more detail, see the X‑Card by John Stavropoulos and the "Lines and Veils" RPG safety toolkit. Many free resources are available online.}
\end{tcolorbox}

\subsection{What Is Fate's Edge?}

Fate's Edge is a **narrative‑first** tabletop roleplaying system where every action carries weight, every choice has consequence, and even failure feeds the story. Dice are not just success meters---they are instruments of fate, weaving opportunity with risk.

\emph{"The road is a ledger. Every broken wheel is a debt, every lit lamp a witness. The only question is: what are you willing to owe?"}

\subsubsection{The Central Question}
\begin{quote}
\emph{What are you willing to risk, and what are you willing to pay, to reshape the world around you?}
\end{quote}

\subsubsection{Tone of Play}
\begin{itemize}
  \item \textbf{Cinematic} -- Scenes breathe, tense exchanges resolve in moments.
  \item \textbf{Consequential} -- A whispered secret can topple a kingdom; a missed roll can spare a life.
  \item \textbf{Collaborative} -- The GM sets stakes; the players drive the fiction.
\end{itemize}

\subsection{The Core Loop in Micro‑Examples}

Before we dive into mechanics, see the rhythm in action:

\begin{examplebox}[Lockpick Under Pressure]
Lyra tries to pick a lock while a guard patrols nearby.\\
\textbf{GM:} "This is \emph{Controlled}, DV 3. On a Miss, the guard hears you."\\
\textbf{Lyra} (Wits 2 + Tinker 2 = 4 dice): rolls 8, 6, 3, 1 → 2 successes (DV 3 → Partial), 1 SB.\\
\textbf{GM:} "The lock clicks open -- but the last pin squeals. I spend 1 SB: the guard stops and tilts his head. You have one beat to hide."
\end{examplebox}

\begin{examplebox}[False Orders, Real Panic]
Kael forges orders to redirect a patrol.\\
\textbf{GM:} "\emph{Controlled}, DV 3. Subterfuge + Wits."\\
\textbf{Kael} (Wits 3 + Subterfuge 2 = 5 dice): rolls 9, 7, 5, 4, 2 → 3 successes, 0 SB. \emph{Clean Success}.\\
\textbf{GM:} "The captain reads, nods, and marches his squad toward the west gate. You"ve bought minutes."
\end{examplebox}

\begin{examplebox}[From Dominant to Desperate]
Sariel fires from behind a pillar (\emph{Dominant} Position).\\
\textbf{Roll:} 6 dice → 10, 9, 7, 6, 3, 1 (1 SB). Clean hit.\\
\textbf{GM spends 1 SB:} "Return fire chips the pillar -- your cover shrinks. Next action is \emph{Controlled}." Later, after two more SB: "Reinforcements arrive -- the scene drops to \emph{Desperate}."
\end{examplebox}

\subsection{Key Concepts}

\subsubsection{Narrative Time}
Measured by story weight, not minutes:
\begin{itemize}
  \item \textbf{A Moment} -- a heartbeat, a strike, a whispered word.
  \item \textbf{Some Time} -- a skirmish, a tense negotiation.
  \item \textbf{Significant Time} -- hours (travel, ritual).
  \item \textbf{Days} -- marches, training, recovery.
\end{itemize}

\subsubsection{Story Beats (SB)}
Every time a player rolls a **1**, the GM gains a Story Beat. SB are narrative currency -- spend them to introduce complications, escalate tension, or shift Position.

\subsubsection{Affinity}
Each culture grants an Affinity -- a narrative edge or metaphysical bond that makes certain actions more reliable (e.g., Aeler stone‑sense, Aelaerem hearth‑luck). Affinities are described in the full Player's Guide.

\subsubsection{Prestige Abilities}
High‑level talents unlocked by mastering cultural or arcane arts. They are narrative milestones as much as mechanical upgrades.

\subsubsection{On‑Screen vs. Off‑Screen}
\begin{itemize}
  \item \textbf{On‑Screen:} followers, allies, companions who stand beside you.
  \item \textbf{Off‑Screen:} assets, titles, informants, safe houses -- resources that act when you spend Boons or during downtime.
\end{itemize}

% --- Revised Boon Section with Fatigue Substitution Rule ---
\subsection{Boons}

Boons are narrative tokens earned primarily by failing rolls with meaningful consequences (see Outcome Matrix). They represent momentum, luck, or a second wind.

\paragraph{Earning Boons}
\begin{itemize}
  \item \textbf{Partial (0 < successes < DV):} Gain 1 Boon.
  \item \textbf{Miss (successes = 0):} Gain 2 Boons (provided the roll was meaningful and generated at least one Story Beat).
  \item \textbf{Bond-driven action:} Once per bond per session, when you act meaningfully on a Mutual Bond and provide an Intricate Description, gain 1 Boon.
  \item \textbf{GM award:} For creative solutions, sacrifices, or outstanding roleplay.
\end{itemize}

\paragraph{Spending Boons}
You may spend Boons on the following effects. Unless otherwise noted, each option costs 1 Boon.
\begin{itemize}
  \item \textbf{Reroll a single die} after seeing the pool (you keep the second result; any new 1s generate additional Story Beats as normal).
  \item \textbf{Activate an on‑screen Asset} (see Assets and Followers).
  \item \textbf{Improve Position} by one step before a roll (e.g., Controlled → Dominant).
  \item \textbf{Power certain Rites or abilities} as described in their text.
  \item \textbf{Convert 2 Boons → 1 XP} (once per session, during downtime; max 2 XP per session via conversion).
\end{itemize}

\paragraph{Substituting Fatigue for Boons}
If you have no Boons remaining (or wish to conserve them), you may pay \textbf{1 Fatigue} in place of 1 Boon to activate any effect that normally costs a Boon. This represents pushing through exhaustion, relying on raw effort, or accepting a minor physical toll. The GM may veto this substitution if the fiction makes it implausible (e.g., activating a delicate Asset while exhausted). Using this option does not change the effect's outcome or generate additional Story Beats; it simply changes the resource spent. \textit{Design note: This rule ensures that players are never completely locked out of Boon‑powered options, but the cost of Fatigue keeps it from being a free substitute.}

\paragraph{Limits}
\begin{itemize}
  \item Maximum 5 Boons held at any time.
  \item At the end of each scene, reduce held Boons to a maximum of 2 (excess are lost).
  \item At most 2 Boons from failures per character per scene (repetition rule: same approach + same stakes cannot award another Boon).
\end{itemize}
% --- PLAYER-MANAGED MODULES (Condensed) ---
\section{Standard Rule: Player-Managed Modules}
\label{sec:player-managed-modules}

Players are the primary stewards of their **character‑facing** modules (Obligation, Corruption, Leash, Asset States). This keeps the table moving and reduces GM overhead.

\subsection{What Players Track}
\begin{itemize}
  \item \textbf{Obligation} (by Patron) -- mark on Invoke/Push/ritual.
  \item \textbf{Corruption Clock} (Cantor) -- mark on Push, Resonant Rite, or GM spending SB on occult actions.
  \item \textbf{Leash} (Summoning) -- mark on spirit harm, unnatural commands, crossing wards.
  \item \textbf{Asset States} (Symbols, tools) -- mark from maintenance or \emph{Crack the Seal}.
\end{itemize}
GM retains all world‑facing clocks and narrative fallout.

\subsection{Procedure}
\begin{itemize}
  \item \textbf{Start of session:} announce current Obligation, Corruption, Leash, and any at‑risk Assets.
  \item \textbf{End of scene:} "Any marks?" -- players update modules; resolve thresholds immediately.
  \item \textbf{Downtime:} apply clears (service, purification, study) -- GM verifies costs.
\end{itemize}

\subsection{Threshold Examples}
\begin{itemize}
  \item \textbf{Obligation Capacity (Spirit+Presence):} Over capacity → +1 Fatigue per segment. Double capacity → clear all Fatigue, +1 Harm, Patron Intrusion.
  \item \textbf{Corruption Full:} gain last‑Patron's boon & burden, reset clock.
  \item \textbf{Leash Full:} spirit acts once to its nature, then departs.
  \item \textbf{Symbol State:} \textsc{Compromised} from \emph{Crack the Seal} -- repair in downtime or pay 1 XP.
\end{itemize}

\subsection{Micro‑Examples}
\begin{itemize}
  \item \textbf{Invoker, two rites:} Vessa Invokes two different Patrons -- marks each Obligation separately. Hitting capacity with Patron A causes Fatigue; Patron B stays below.
  \item \textbf{Cantor Pushes:} Jorel Pushes a Song -- marks +1 Corruption. That fills his clock; he immediately gains last‑Patron's boon/burden, then resets to 0.
  \item \textbf{Summoner Clash:} Kestra's elemental takes Harm and crosses a ward -- she ticks Leash twice. On fill, the elemental flares once and departs.
\end{itemize}

\subsection{Final Note}
The GM remains the arbiter of Patron intrusions, faction moves, and major narrative consequences. Player‑managed modules keep the game fast and transparent, but the story's weight belongs to the whole table.