% Fate's Edge SRD --- Equipment (Section 06)

\section{Equipment}

\subsection{Weapons by Weight Class}

A weapon's weight determines its Harm, its dice modifier at different ranges, and its tempo. Harm is applied directly when an attack hits; the dice modifier affects the attack roll, not damage.

\begin{tabular}{@{}lccccl@{}}
\toprule
\textbf{Weight} & \textbf{Harm} & \textbf{Close} & \textbf{Near} & \textbf{Far} & \textbf{Tempo} \\
\midrule
Light   & 1 & +2d & +1d & --- & Fast \\
Medium  & 2 & +1d & +2d & --- & Standard \\
Heavy   & 3 & –1d & +3d & +1d & Slow \\
\bottomrule
\end{tabular}

\subsubsection{Notes}
\begin{itemize}
  \item \textbf{Fast tempo:} You may move and attack in the same turn (Move + Action).
  \item \textbf{Standard tempo:} You may move or attack in a turn, not both.
  \item \textbf{Slow tempo:} Attacking requires a Set action (or a full turn action) before the attack; you cannot move and attack in the same turn unless a talent says otherwise.
  \item \textbf{Heavy weapons at Close:} Suffer –1d due to limited maneuvering room.
  \item A weapon cannot attack beyond its maximum listed range unless a talent or special ammunition extends it.
\end{itemize}

\subsection{Ranged Weapons and Ammunition}

Ranged weapons follow the same weight and tempo rules. Ammunition is tracked abstractly as part of Supply (see Section~\ref{sec:travel-framework}). A character may run low on ammunition when the GM spends SB on resource strain or during extended travel.

\subsection{Armor}

Armor does not reduce Harm directly. Instead, it converts incoming Harm into Fatigue. See Section~\ref{sec:core-mechanic} for the full conversion procedure.

\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Armor Type} & \textbf{Conversion} & \textbf{Notes} \\
\midrule
Light   & 1 Harm $\rightarrow$ 1 Fatigue & –1d to physical actions (optional) \\
Medium  & 2 Harm $\rightarrow$ 1 Fatigue & No penalty \\
Heavy   & 3 Harm $\rightarrow$ 2 Fatigue & –2d to physical actions; cannot sprint \\
\bottomrule
\end{tabular}

\subsubsection{Armor Conversion Examples}
\begin{itemize}
  \item Light armor vs Harm 1: take 1 Fatigue. No Harm.
  \item Medium armor vs Harm 2: take 1 Fatigue. No Harm.
  \item Heavy armor vs Harm 3: take 2 Fatigue. No Harm.
  \item Minimum 1 Fatigue per hit, even if conversion would give 0.
\end{itemize}

\subsection{Shields (Optional)}

Shields are worn in the off-hand. Using a shield restricts weapon size to Light or Medium.

\begin{tabular}{@{}lcl@{}}
\toprule
\textbf{Shield} & \textbf{Benefit} & \textbf{Tradeoff} \\
\midrule
Buckler    & +1d to Defend actions once per scene & Off-hand occupied; –1d to ranged attacks \\
Heater     & +1d to Defend actions; 1 Harm $\rightarrow$ Fatigue per scene & Off-hand occupied; –1d to ranged attacks \\
Pavise     & Plant to create heavy cover (one direction) & Bulky; cannot move while planted \\
\bottomrule
\end{tabular}

\subsection{Mundane Gear}

Most gear is abstracted. When you need a specific item (rope, lockpicks, torch), assume you have it unless the fiction says otherwise or the GM spends SB to declare a shortage.

The following items are tracked as part of \textbf{Supply} (see Section~\ref{sec:travel-framework}):

\begin{multicols}{2}
\begin{itemize}
  \item Rations (1 day)
  \item Torches (3)
  \item Waterskin
  \item Rope (10m)
  \item Tinderbox
  \item Bedroll
  \item Crowbar
  \item Hammer and spikes
  \item Sack (large)
  \item Grappling hook
  \item Lantern and oil
  \item Healer's kit (3 uses)
\end{itemize}
\end{multicols}

\subsection{Equipment Condition}

All gear can become \textbf{Neglected} or \textbf{Compromised}.

\begin{itemize}
  \item \textbf{Neglected:} Weapons suffer –1 die on attack rolls; armor conversion worsens by one step (e.g., Light behaves as Unarmored). Neglected gear is usable but unreliable.
  \item \textbf{Compromised:} Weapons suffer –1 die on the first attack each round; armor provides no conversion (all Harm applies directly). Compromised gear is barely functional.
\end{itemize}

\subsubsection{Repair}
\begin{itemize}
  \item \textbf{Neglected $\rightarrow$ Functional:} Requires a short rest (1 hour) and access to basic tools.
  \item \textbf{Compromised $\rightarrow$ Neglected:} Requires a scene of dedicated work (Craft test, DV 3) or a visit to a smith.
  \item \textbf{Compromised $\rightarrow$ Functional:} Requires significant downtime (a full day) and a successful Craft test (DV 4) or a professional smith.
  \item Enchanted items require the same repairs; enchantments are inactive while the item is Compromised.
\end{itemize}

\subsection{Magical Items (Enchantments)}

Enchantments are permanent magical modifications treated as Minor or Standard Assets (see Section~\ref{sec:followers-assets}). Each enchantment costs XP and counts toward the item's enchantment limit.

\subsubsection{Enchantment Limits}
\begin{itemize}
  \item Light item: maximum 1 enchantment.
  \item Medium item: maximum 2 enchantments.
  \item Heavy item: maximum 3 enchantments.
\end{itemize}

\subsubsection{Sample Enchantments}
\begin{itemize}
  \item \textbf{Keen Edge (2 XP):} +1 die when attacking armored targets.
  \item \textbf{Flaming Blade (4 XP):} +1 Harm against creatures vulnerable to fire; +1 Effect against cold-based foes.
  \item \textbf{Soulfire Weapon (6 XP):} Ignores 1 point of armor or resistance.
  \item \textbf{Shadowweave Armor (2 XP):} +1 die to Stealth rolls while moving.
  \item \textbf{Runed Plate (4 XP):} Reduce magical backlash severity by one step (e.g., Moderate $\rightarrow$ Minor).
  \item \textbf{Boots of Sure Step (3 XP):} Ignore difficult terrain penalties.
\end{itemize}

\subsubsection{Removing an Enchantment}
An enchantment can be removed with a ritual (Wits + Arcana, DV 4) costing 1 XP per enchantment level. The item is not damaged.

\subsection{Weapon and Armor Acquisition}

\begin{itemize}
  \item Mundane weapons and armor are acquired through play (found, looted, purchased with Supply, or granted as part of character background). No XP cost.
  \item Masterwork or enchanted items require XP expenditure (as Assets) or narrative quests.
  \item A character begins play with one Light or Medium weapon and one set of Light or Medium armor appropriate to their background.
\end{itemize}

