% Fate's Edge SRD --- Action Economy (Section 07)

\section{Action Economy}

\subsection{Turn Structure}

Each character takes \textbf{1 Action} and \textbf{1 Move} on their turn. Actions and Moves may be taken in any order. A character may use a Boon to re-roll their action at the expense of their Move if they still have it available.

\subsection{Move}

\begin{itemize}
  \item Shift one range band: Close $\leftrightarrow$ Near $\leftrightarrow$ Far.
  \item Reposition within the same range band (e.g., from behind one tree to another). This does not change your range band but may affect cover or line of sight.
  \item \textbf{Stand:} Use your Move to stand from prone.
\end{itemize}

\subsection{Actions}

\subsubsection{Attack}
Make a melee or ranged attack versus DV set by the GM and fiction. Ranged attacks made while within Close range of an enemy are Desperate (re-roll one success).

\subsubsection{Disengage}
Action (not a Move). You prepare to retreat safely. After taking Disengage, your Move this turn does not provoke opportunity attacks. You cannot also Attack this turn.

\subsubsection{Dash}
Action. Move two range bands (e.g., Close $\rightarrow$ Far). Your next defense roll is Desperate.

\subsubsection{Charge (Full Turn Action)}
Move up to two range bands toward an enemy and make one melee attack. Costs 1 Fatigue. Your next defense is Desperate. Consumes both your Action and Move. Requires open ground (no difficult terrain). If your target is behind full cover, you cannot complete the Charge --- you may move up to the cover and end your turn there; the attack is lost.

\subsubsection{Skirmish (Full Turn Action)}
Move up to two range bands away from an enemy and make one ranged attack. Costs 1 Fatigue. Your next defense is Desperate. Consumes both Action and Move. Requires open ground. If you start your turn hidden, you may Skirmish: leave cover, move, shoot, and may end your turn in new cover if available. You are exposed during movement and the shot; opportunity attacks apply normally.

\subsubsection{Observe / Change Position}
Take a beat to read the field or set angles. Gain +1 Position for one action this turn (e.g., Controlled $\rightarrow$ Dominant). Limit once per turn; cannot exceed Dominant.

\subsubsection{Setup (Teamwork)}
Create advantage for an ally. On success, grant their next action +1 Position or step up Effect (GM's call).

\subsubsection{Assist (Teamwork)}
Spend 1 Boon to give an ally +1 die on their current roll. You share appropriate risk or consequence.

\subsubsection{Protect}
Adopt a guarding stance or body-block. Choose a nearby ally. Until your next turn, you may intercept one hit targeting them and roll to resist it. On success, reduce or negate Harm; you take any fallout the GM assigns.

\subsubsection{Interact}
Lift, pull, flip a lever, shove a foe, break an object, apply a poultice, reload, draw or stow a weapon, hide, or search. GM sets DV and Effect.

\subsubsection{Defend}
Spend your Action or Move to enter a defensive stance. Until your next turn, you count as Defending. When resisting any attack or effect, improve your Position by one step (or gain +1 die if already Dominant).

\subsubsection{Free Actions}
Short shouts, dropping an item, a quick glance. Longer or tactical assessments require Observe / Change Position or Interact.

\subsection{Reactions (Out of Turn)}

\begin{itemize}
  \item \textbf{Opportunity Attack:} When an enemy leaves your Close range without using Disengage, you may make one free melee attack (Controlled Position, no action cost). This does not consume your action for the round.
  \item \textbf{Protection} may trigger when an ally is hit and you are in position.
  \item Class and asset reactions fire as written (e.g., counter-runes, ripostes).
\end{itemize}

Actions that provoke opportunity attacks when performed while engaged in Close with an enemy:
\begin{itemize}
  \item Making a ranged attack (bow, crossbow, thrown weapon).
  \item Casting a spell that does not have the \texttt{[TOUCH]} or \texttt{[WARD]} tag.
\end{itemize}

\subsection{Sidestep Strike}

As part of a melee attack action, you may move one range band (Close $\leftrightarrow$ Near or Near $\leftrightarrow$ Far) without using your Move action for the turn. If you do so, your Position for that attack worsens by one step (e.g., Controlled becomes Desperate, Dominant becomes Controlled). You cannot also use your regular Move action in the same turn.

\subsection{Shove}

Action. Opposed Body + Athletics vs. target's Body + Athletics. On success, push the target back one range band (Close $\rightarrow$ Near). The target suffers --1 Position on their next action.

\subsection{Grapple}

Action. Opposed Body + Melee vs. target's Body + Athletics. On success, both characters are engaged (cannot move apart). The grappled target suffers --1 die on all actions. Breaking free requires a new opposed roll as an action.

\subsection{Hiding and Cover}

\begin{itemize}
  \item To hide, a character must be behind \textbf{full cover} (e.g., a large tree, a wall) or in heavy concealment (dense fog, darkness).
  \item Take the \textbf{Interact} action to hide, rolling Stealth (or appropriate skill) vs. the Perception of any observers.
  \item Once hidden, you are not a valid target for attacks unless an attacker first uses an action to \textbf{Search} (Perception vs. your Stealth) or you reveal yourself by attacking or moving into the open.
  \item If you move out of cover to attack, you are no longer hidden for the duration of the attack and may be subject to opportunity attacks if you leave an enemy's Close range.
\end{itemize}

\subsection{Ready and Overwatch}

Players may hold an action with a clear trigger (e.g., ``I shoot the first enemy who comes through that door''). If the trigger occurs, resolve the held action before the next exchange starts.

\subsection{Refresh Windows}

\begin{itemize}
  \item \textbf{Once per scene} effects reset at scene end.
  \item \textbf{Once per exchange} effects reset at the next beat.
\end{itemize}

\subsection{Summary Table of Actions}

\begin{tabular}{@{}lp{5cm}p{5cm}@{}}
\toprule
\textbf{Action} & \textbf{Cost} & \textbf{Effect} \\
\midrule
Attack & 1 Action & Make melee or ranged attack. \\
Disengage & 1 Action & Move does not provoke opportunity attacks; cannot Attack. \\
Dash & 1 Action & Move two bands; next defense Desperate. \\
Charge & Full turn (Action + Move), 1 Fatigue & Move two bands toward enemy + melee attack; next defense Desperate. \\
Skirmish & Full turn (Action + Move), 1 Fatigue & Move two bands away + ranged attack; next defense Desperate. \\
Observe & 1 Action & +1 Position on next action this turn. \\
Setup & 1 Action & Grant ally +1 Position or +1 Effect. \\
Assist & 1 Boon & Grant ally +1 die on current roll. \\
Protect & 1 Action & Intercept one hit on nearby ally. \\
Interact & 1 Action & Manipulate environment, hide, search, etc. \\
Defend & 1 Action or Move & Defensive stance; +1 Position or +1 die to resist. \\
\bottomrule
\end{tabular}

