% --- Fate's Edge SRD — Section 7: Complications ---
\section{Complications}

All actions in Fate's Edge carry risk. Complications are the narrative, mechanical, and environmental consequences that arise from failure, pressure, or the GM spending Story Beats.

\subsection{Story Beats (SB)}
Story Beats are the GM’s primary currency for introducing complications. Each die result of \textbf{1} generates 1 SB. Re-rolling a 1 does not erase its SB; if the re-rolled die also shows 1, it generates additional SB.

\paragraph{Spending SB}
The GM spends SB to introduce complications. Use the following menu as a guideline:
\begin{itemize}
  \item \textbf{1 SB}: Minor pressure – noise, trace, tick a 4/6 clock +1, worsen Position by one step.
  \item \textbf{2 SB}: Moderate setback – alarm raised, lose cover/advantage, a lesser foe appears, split the party briefly.
  \item \textbf{3 SB}: Serious trouble – reinforcements arrive, key gear breaks (temporarily), a major twist shifts the scene.
  \item \textbf{4+ SB}: Scene-shaping turn – a trap springs, authority arrives, the environment changes dramatically, a Patron manifests an omen.
\end{itemize}

\paragraph{Limits on SB per Scene}
\begin{itemize}
  \item Base SB budget: 4 + highest character Tier (e.g., Tier I = 5, Tier III = 7).
  \item Maximum 12 SB for standard scenes, 16 SB for climactic scenes.
  \item Unspent SB may carry over between scenes (GM discretion, max 4).
\end{itemize}

\subsection{Boons}
Boons are the players’ resource for mitigating complications or turning the tide.

\paragraph{Earning Boons}
\begin{itemize}
  \item \textbf{Miss (0 successes):} Gain 2 Boons (provided the roll was meaningful and SB was generated).
  \item \textbf{Partial (0 < successes < DV):} Gain 1 Boon.
  \item \textbf{Bond-driven action:} Once per bond per session, when you act in a way that meaningfully engages a bond and you provide an Intricate Description, gain 1 Boon.
  \item \textbf{GM award:} For creative solutions, sacrifices, or outstanding roleplay.
\end{itemize}

\paragraph{Spending Boons}
\begin{itemize}
  \item Re-roll a single die in any pool (after seeing the result).
  \item Activate an on-screen Asset (1 Boon).
  \item Improve Position by one step before a roll.
  \item Power certain Rites or abilities.
  \item Convert 2 Boons → 1 XP (once per session, during downtime; max 2 XP per session).
\end{itemize}

\paragraph{Limits on Boons}
\begin{itemize}
  \item Hold up to 5 Boons. At scene end, reduce to 2 (excess lost).
  \item Max 2 Boons earned from failures per character per scene.
\end{itemize}

\subsection{Clocks as Complications}
Clocks track accumulating threats or progress. They are a visible way to manage pressure.

\paragraph{When to Use a Clock}
\begin{itemize}
  \item \textbf{Threat Clock (4–8 segments):} Fills with enemy action, environmental decay, or social collapse. When full, a defined disaster occurs.
  \item \textbf{Progress Clock (4–8 segments):} Fills with player successes to achieve a goal (e.g., escape, research, negotiation).
  \item \textbf{Contest Clocks:} Two clocks race – first to fill wins.
\end{itemize}

\paragraph{Ticking Clocks}
Clocks advance when:
\begin{itemize}
  \item The GM spends SB (e.g., 1 SB = tick a relevant clock +1).
  \item A failed roll directly relates to the threat (GM discretion).
  \item A scene transition or scripted event occurs.
\end{itemize}

\paragraph{Example Clocks}
\begin{itemize}
  \item \textbf{Guards Incoming [6]} – ticks on noise or missed stealth. On fill, a patrol arrives.
  \item \textbf{Collapsing Bridge [4]} – ticks on failed Athletics or heavy impacts. On fill, bridge falls.
  \item \textbf{Patron’s Patience [8]} – ticks each time Obligation is ignored. On fill, Patron demands service.
\end{itemize}

\subsection{Hazards (Environmental, Social, Magical)}
Hazards are ongoing complications that require action to avoid or mitigate.

\paragraph{Environmental Hazards}
\begin{itemize}
  \item \textbf{Active:} Fire spreading, flooding, toxic gas – represented by a Threat Clock that ticks each round or on SB spends.
  \item \textbf{Passive:} Difficult terrain, darkness, extreme heat/cold – impose Position penalties or DV modifiers.
\end{itemize}

\paragraph{Social Hazards}
\begin{itemize}
  \item \textbf{Rumors:} A clock that advances when the party fails social rolls; when full, a faction turns hostile.
  \item \textbf{Scrutiny:} Worsens Position for stealth or deception.
  \item \textbf{Reputation Damage:} Future social rolls start at worse Position.
\end{itemize}

\paragraph{Magical Hazards}
\begin{itemize}
  \item \textbf{Backlash Zone:} An area where free casting automatically generates +1 SB per spell.
  \item \textbf{Ward / Seal:} Requires a roll to pass; failure triggers a consequence (harm, alarm, etc.).
  \item \textbf{Leaking Veil:} Spirits or fey occasionally manifest spontaneously (GM spends SB to introduce).
\end{itemize}

\subsection{Fear Effects}
When a character escalates on the Fear Track (Shaken → Frightened → Panicked), roll or choose an effect from the table below. Apply to NPCs by default; PCs may adopt as narrative guidance.

\begin{center}
\begin{longtable}{>{\bfseries}r l l}
\toprule
d10 & Effect & Magic Tags \\
\midrule
1 & \textbf{Freeze}: Cannot act this round, staring or trembling. & Silence, Stasis \\
2 & \textbf{Flee}: Must move at full speed away from the source of Fear. & Movement, Wind \\
3 & \textbf{Drop}: Drops what they are holding. & Disarm, Break \\
4 & \textbf{Beg}: Pleads or bargains incoherently. & Compulsion, Voice \\
5 & \textbf{Hide}: Seeks cover, concealment, or allies. & Shadow, Illusion \\
6 & \textbf{Attack in Panic}: Lashes out wildly at the nearest target. & Rage, Fire \\
7 & \textbf{Blunder}: Stumbles into danger (trap, hazard, off balance). & Chaos, Trickery \\
8 & \textbf{Obey}: Instinctively follows a simple command from the fear-causer. & Command, Charm \\
9 & \textbf{Break Down}: Sobs, prays, or becomes useless until aided. & Curse, Despair \\
10 & \textbf{Catatonia}: Becomes unresponsive, requiring intervention. & Sleep, Dream \\
\bottomrule
\end{longtable}
\end{center}

\paragraph{Fear Track Escalation}
\begin{itemize}
  \item \textbf{Shaken:} Apply minor effects (hesitation, –1 die, startled).
  \item \textbf{Frightened:} Roll normally on the table.
  \item \textbf{Panicked:} Apply severe or exaggerated results (e.g., 2 = reckless flight into danger, 6 = attacks allies).
\end{itemize}

\subsection{Integrating Complications with SB and Boons}
\begin{itemize}
  \item SB and Boons create a push-pull: SB gives the GM tools to escalate; Boons give players tools to recover or resist.
  \item A well-spent SB should change the situation, not just punish. Offer the player a Devil’s Bargain: “I’ll spend 2 SB to have the guard turn around – but you’ll owe a favor.”
  \item Clocks make threat visible; when a clock ticks, describe what changes in the fiction.
\end{itemize}

\subsection{Example: Complication in Action}
\emph{Lyra misses a stealth roll (0 successes) while sneaking past a guard post. She gains 2 Boons. The GM spends 1 SB: “Your foot scrapes a loose stone – the guard looks up.” (Tick Guards Alert [4] +1). Later, the GM spends another 2 SB: “The guard calls for backup – two more arrive. New clock: Reinforcements [6].” Lyra uses 1 Boon to re-roll a failed athletics test to climb a wall, escaping just as the clock reaches 4.}