% --- Fate's Edge SRD — Section 3: Combat System ---
% Include this file from your main .tex with: \input{03-combat.tex}

\section{Combat System}

\subsection{Core Philosophy}
Combat in Fate's Edge is not a separate mini-game; it is simply conflict under sharper focus. 
It uses the same dice pool system and SB economy as all other actions. 
The rules emphasize narrative consequence, positional play, and risk management.

\subsection{Structure of Combat}
\begin{itemize}
  \item \textbf{Rounds:} Each round represents a few seconds of action.
  \item \textbf{Turns:} Each participant takes one significant action per round.
  \item \textbf{Scenes:} A battle is one scene unless the fiction dictates otherwise.
\end{itemize}

\subsection{Taking Action}
On your turn:
\begin{enumerate}
  \item \textbf{Set Position:} The GM decides if you are Controlled, Risky, or Desperate.
  \item \textbf{Build Pool:} Attribute + Skill (+ gear, + assists, +1 from Imbuement if active).
  \item \textbf{Roll:} Each 6+ is a success. Each 1 generates SB.
  \item \textbf{Resolve:} Compare successes to DV; GM applies Position, Effect, and SB costs.
\end{enumerate}

\subsection{Position \& Effect}
\begin{description}[leftmargin=1.5em]
  \item[Controlled] You act from safety or advantage. Failure still leaves you options.
  \item[Risky] Standard case. Failure has teeth, but not ruin.
  \item[Desperate] High stakes. Failure is severe; success may bring extra XP (mark Desperate use).
\end{description}

\paragraph{Effect}
\begin{itemize}
  \item \textbf{Limited:} Scratch or slow progress.
  \item \textbf{Standard:} Expected impact (downing a guard, disabling a lock).
  \item \textbf{Great:} Overwhelming impact, bigger than expected.
\end{itemize}

% =========================
% HEALTH, FATIGUE, & HARM
% =========================
\section{Health, Fatigue, \& Harm}
\label{sec:health-fatigue-harm-rev}

\subsection*{Tracks \& Caps}
\begin{itemize}
  \item \textbf{Fatigue Track:} boxes equal to \textbf{Body}.
  \item \textbf{Harm Levels:} Harm~1 (minor), Harm~2 (serious), Harm~3 (critical/incapacitated).
\end{itemize}

\subsection*{Fatigue $\rightarrow$ Harm Conversion}
When Fatigue fills:
\begin{enumerate}
  \item Increase Harm by 1 level.
  \item Clear all Fatigue.
\end{enumerate}
This can occur multiple times per scene.

\subsection*{Taking Fatigue}
Mark Fatigue for strain, exertion, travel, magic costs, or \S\ref{sec:obligation-overflow-rev} overflow.

\subsection*{Recovering Fatigue}
\begin{itemize}
  \item \textbf{Short Rest:} remove 2 Fatigue.
  \item \textbf{Full Night:} remove all Fatigue.
\end{itemize}
Harm is not removed by Fatigue recovery.

\subsection*{Mitigation}
\begin{itemize}
  \item \textbf{Soak/Ward:} Reduce Fatigue by 1–2 before marking if shielded.
  \item \textbf{Convert:} Some effects turn Harm~1 into 2 Fatigue.
\end{itemize}

\subsection{Damage \& Consequences}
\begin{itemize}
  \item \textbf{Harm 1:} -1 die to related actions.
  \item \textbf{Harm 2:} -1 die to most actions until treated.
  \item \textbf{Harm 3:} Incapacitated or dying.
\end{itemize}
Harm may be resisted (roll Attribute vs DV~3; 1s generate SB).

% =========================
% WEAPONS & ARMOR
% =========================
\section{Weapons \& Armor}
\label{sec:weapons-armor}

\subsection*{Weapons}
Purchased by weight class. Provides modifiers by range.
\begin{itemize}
  \item \textbf{Light (4 XP):} Melee +2 Close, +1 Near. Ranged +1 Near. Fast, no penalties.
  \item \textbf{Medium (8 XP):} Melee +1 Close, +2 Near. Ranged +2 Near, +1 Far. Must \emph{Set} once per scene or roll -1 die first attack.
  \item \textbf{Heavy (12 XP):} Melee -1 Close, +3 Near. Ranged +1 Near, +3 Far. Must \emph{Set} once per scene or roll -2 dice first attack.
\end{itemize}

\subsection*{Armor}
Armor converts Harm into Fatigue.
\begin{itemize}
  \item \textbf{Light (4 XP):} Up to 1 Harm $\rightarrow$ 1 Fatigue.
  \item \textbf{Medium (8 XP):} Up to 2 Harm $\rightarrow$ 1 Fatigue. -1 die physical skills.
  \item \textbf{Heavy (12 XP):} Up to 3 Harm $\rightarrow$ 2 Fatigue. -2 dice physical skills, no sprint in rough terrain.
\end{itemize}

\subsection*{Condition}
\begin{itemize}
  \item \textbf{Neglected:} Conversion reduced by 1 step; weapons roll -1 die.
  \item \textbf{Compromised:} No conversion; weapons roll -1 die first attack each exchange.
\end{itemize}

% =========================
% RANGED COMBAT
% =========================
\section{Ranged Combat}
\label{sec:ranged}

\subsection*{Range Bands}
\begin{description}[leftmargin=1.5em]
  \item[Close:] Room-length, melee-adjacent.
  \item[Near:] Street, courtyard, glade.
  \item[Far:] Field, battlement, valley.
\end{description}

\subsection*{Firing Profiles}
\begin{description}[leftmargin=1.5em]
  \item[Fast:] (knives, slings, shortbows) — Move and Shoot. Reload is incidental. Risky at Close.
  \item[Standard:] (longbows, crossbows) — Move or Shoot. Reload takes part of turn. Aim for +1 die or Effect.
  \item[Slow:] (greatbows, arbalests, siege) — Requires Set/Brace. Reload full turn. Cannot Move and Shoot same round.
\end{description}

\subsection*{Options}
\begin{itemize}
  \item \textbf{Aim:} +1 die or Effect next shot (lost if disrupted).
  \item \textbf{Volley:} Extra ammo = +1 die (max +2).
  \item \textbf{Suppress:} Zone fire; foes act at -1 die/Limited Effect.
  \item \textbf{Overwatch:} Hold, fire Risky shot when target crosses line.
\end{itemize}

\subsection*{Cover \& Close}
\begin{itemize}
  \item Light cover: +1 DV or -1 die to attacker.
  \item Heavy cover: +2 DV or -2 dice.
  \item Obscured (fog, night): Position worsens by 1 step.
  \item Shooting while Engaged: Risky (Fast) or Desperate (Standard/Slow).
\end{itemize}

\subsection*{Weapon Tags (defaults)}
\begin{itemize}
  \item Light ranged = Fast.
  \item Medium ranged = Standard.
  \item Heavy ranged = Slow.
\end{itemize}

% =========================
% TEAMWORK & SB
% =========================
\subsection{Teamwork}
\begin{itemize}
  \item \textbf{Assist:} Spend 1 Stress or Boon for +1 die (max +3).
  \item \textbf{Setup:} Roll to improve ally’s Position or Effect.
  \item \textbf{Protect:} Take harm meant for another.
\end{itemize}

\subsection{GM Guidance for SB}
\begin{itemize}
  \item \textbf{1 SB:} Reinforce cover, minor injury, new foe.
  \item \textbf{2 SB:} Reinforcements, gear breaks, +1 die foe.
  \item \textbf{3 SB:} Rite unleashed, terrain shifts, ally endangered.
  \item \textbf{4+ SB:} Scene twist—fire spreads, floor collapses, omen manifests.
\end{itemize}

\subsection{Combat and Magic}
\begin{itemize}
  \item \textbf{Casting:} Weave (1 action), Cast (1 action). Requires Caster’s Gift.
  \item \textbf{Rites:} Invoke in 1 action. Push = +1 Obligation.
  \item \textbf{Invokers:} Rituals take time, or Crack the Seal (instant, costs Obligation).
  \item \textbf{Imbuements:} 1 action once per scene; +1 to one Weapon + one Skill.
\end{itemize}

\subsection{Worked Example}
\emph{Kael swings his Imbued blade at a cultist (DV 2). He rolls 5 dice: 9, 7, 5, 2, 1.}
\begin{itemize}
  \item Successes = 3 (hit), SB = 1.
  \item GM: cultist falls.
  \item GM spends 1 SB: ``Blood sprays across the sigil—energy flares, the summoning accelerates.''
\end{itemize}
