% =========================
% Fate's Edge SRD — Section 09: Rites, Invokers, and Symbols
% Include from main with: \input{patrons/09-rites-invokers.tex}
% Requires: \usepackage{float} for [H] tables
% =========================

\section{Rites, Invokers, and Symbols}
\label{sec:rites}

Magic in \textbf{Fate's Edge} expresses through three intertwined practices: \textbf{Rites} (oathbound authority), \textbf{Invocations} (symbolic ritual), and \textbf{Patron Pacts} (gifts and obligations). The rules below emphasize fiction-first play: consequences are Story Beats (SB) that prompt twists; numbers follow the story.

\subsection{Rites and Patrons (Runekeepers)}
\label{subsec:runekeepers}
Characters who bind themselves to a \emph{single} Patron and study that Patron's \textbf{Codex} are \textbf{Runekeepers}. Their magic is structured, immediate, and tied to service.

\begin{itemize}
  \item \textbf{One-Patron Rule.} A Runekeeper may be bound to \emph{only one} Patron at a time. This sharpens identity and keeps Obligation on a single ledger.
  \item \textbf{Thiasos (Familiar).} A circle, retinue, or emissary that grounds the pact in fiction. Required to access \emph{Patron's Gift}.
  \item \textbf{Codex.} The Patron's corpus of rites and precedents. Grants access to the Patron's Rites.
  \item \textbf{Invoke Rites.} A Runekeeper may Invoke a known Rite from their Patron as a \textbf{1 action} effect. On completion, mark \textbf{+1 Obligation} to that Patron. You may \emph{Push It} once per scene to amplify the effect, marking \textbf{+1 additional Obligation}.
\end{itemize}

\subsection*{Rites Difficulty Value (DV)}

The DV to cast a Rite is:

\[
\text{DV} = \max\!\big(\text{Obligation Cost} - \text{Spirit}, \, \text{Tier}\big)
\]

\begin{itemize}
  \item \textbf{Obligation Cost:} The Rite’s listed cost in Obligation segments.
  \item \textbf{Spirit:} The caster’s Spirit attribute, reducing the effective cost.
  \item \textbf{Tier:} The Rite’s minimum difficulty based on its scope or level.
\end{itemize}

\paragraph{Interpretation.} The Obligation burden sets the baseline challenge, but personal Spirit mitigates it. No Rite can be easier than its Tier.

\section{Obligation Capacity}

A character’s \textbf{Obligation Capacity} equals Spirit + Presence.  
Track total Obligation segments across all Patrons (or Symbols, for Invokers).

\begin{itemize}
  \item \textbf{Exceeding Capacity:} For each segment above Capacity, mark 1 Fatigue. The character cannot Invoke Rites or perform rituals until Obligation is reduced below Capacity.
  \item \textbf{Resolution:} Reduce Obligation through Downtime service, Patron tasks, ritual cleansing, or story resolution.
\end{itemize}

\textbf{Example:} Spirit~2 + Presence~3 = Capacity 5.  
6 segments → Fatigue~1.  
7 segments → Fatigue~2.  
10 segments → Harm~1.  
11 segments → Harm~2.


\centering
\caption{Universal Push It Costs}
\begin{longtable}{|l|l|}
\hline
\textbf{Cost Component} & \textbf{Effect} \\
\hline
+1 SB & Escalate effect immediately \\
+1 Fatigue & Immediate physical/mental strain \\
+1 Corruption Clock Segment & Long-term Patron influence (unless otherwise specified) \\
GM spends 1 SB & Thematic complication (unless otherwise specified) \\
\hline
\end{longtable}


Note: Some talents, Rites, or magical paths may specify alternative corruption costs or additional consequences for Push It actions. When explicitly stated, those specific rules override the universal costs.


\subsection{Invokers and Symbols}
\label{subsec:invokers}
Invokers relate to Patrons through consecrated \textbf{Symbols}: physical tokens that anchor names and permissions.

\begin{itemize}
  \item \textbf{Symbols (Minor Asset).} Each Symbol is keyed to one Patron; cost \textbf{4 XP}. You may own Symbols of different Patrons (one Symbol per Patron).
  \item \textbf{Ritual Invocation.} Display the Symbol and perform the Rite as a \emph{ritual} (DV + 1 rounds). Completion always marks \textbf{+1 Obligation} on that Rite's ledger.
  \item \textbf{Crack the Seal.} As part of an Invoker Rite, you may resolve the effect instantly by setting the Symbol to \emph{Compromised} and marking \textbf{+2 Obligation} (\textbf{+3} if High-Power). The Keeper may spend 1 on-theme SB immediately. The asset remains but is inert until restored.
  \item \textbf{Restore a Symbol.} 1 downtime action and a fitting test (DV 3 or by fiction). Success: \emph{Maintained}; shaky: returns \emph{Neglected}. Or spend \textbf{1 XP} to fully restore.
  \item \textbf{Display Requirement.} Symbols must be openly displayed for rituals. Hidden Symbols do not function.
\end{itemize}

\subsection{Casting and Free-Form Magic}
\label{subsec:casting}
Improvised casting is possible with the \textbf{Spellcraft} Talent (\textbf{6 XP}). It is a \emph{backup toolkit}:
\begin{itemize}
  \item Small, local effects (typ. DV 2--3), fiction-first, colored by Elements and locus.
  \item Heavy control effects such as \texttt{[WARD]}, \texttt{[BANISH]}, or \texttt{[UNWARD]} require a printed Talent, Rite, or Spell result.
\end{itemize}

\subsection{Patron's Gift (Imbuements)}
\label{subsec:patrons-gift}
The pact may mark a devotee's tools with a short-lived boon aligned to the Patron's domain.

\paragraph{Requirements.}
\textbf{Thiasos (Familiar)} is required. Invoking the Gift costs \textbf{1 Boon}. A Codex is \emph{not} required for the Gift.

\paragraph{Activation and Duration.}
\begin{itemize}
  \item \textbf{Action:} 1 action to activate; \textbf{1/scene}.
  \item \textbf{Duration:} Scene. \emph{Push It:} extend for one additional scene by marking \textbf{+1 Obligation} to that Patron (max one Push per scene).
  \item \textbf{Range:} Touch (you must handle the item).
  \item \textbf{Stacking:} Gifts from the \emph{same Patron} do not stack; take the best active version. Dice bonuses respect the table's \textbf{+3 dice cap}.
\end{itemize}

\paragraph{Effect.}
Choose one held item you or an ally carries. Until scene end it grants:
\begin{itemize}
  \item \textbf{+1 Melee} (the item counts as a magical weapon), and
  \item \textbf{+1 Thematic} (a \emph{+1 die} to a fixed Skill tied to your Patron; see Table~\ref{tab:gift-thematic-map}). Apply only when the fiction clearly fits the Patron's sphere and how the item is used.
\end{itemize}

% !TEX root = srd_main.tex
% SRD Insert: Clearing Obligation Segments (Condensed SRD Version)

\section{Clearing Obligation Segments (Condensed)}\label{sec:obligation-clearing-srd}
\index{Obligation}\index{Patrons}\index{Acts of Service}\index{Boon}\index{Focused Devotion}

Obligation tracks the narrative weight of a Patron’s attention. To reduce it, characters must act in ways that reaffirm their relevance.

\subsection*{Off-Screen Clearing (Primary)}
\begin{itemize}
\item \textbf{Act of Service.} Clear 1 segment by performing a thematically resonant act of service to the Patron.
\item \textbf{Requirement.} The player provides an \emph{Intricate Description} of the act.\index{Acts of Service}
\item \textbf{Scope.} Not arduous, but meaningful and aligned to the Patron’s sphere.
\end{itemize}

\subsection*{On-Screen Clearing (Optional)}
\paragraph{Talent: Focused Devotion (3–5 XP).}\index{Focused Devotion}
Once per scene, spend 1 Boon to clear 1 segment of Obligation for a Patron. Cannot be used on segments accrued in the same scene.

\subsection*{Narrative Implications}
\begin{itemize}
\item \textbf{Agency.} Players choose how service manifests.
\item \textbf{Collaboration.} GM approves sufficiency of the act.
\item \textbf{Dynamics.} Patron bonds remain active and evolving.
\item \textbf{Capping.} Prevents snowball clearing during a scene; off-screen remains the main method.
\end{itemize}

\paragraph{Runekeeper Clarification.}
A Runekeeper (one Patron + Codex) may Invoke Rites on-screen and use Patron's Gift if they also possess \textbf{Thiasos (Familiar)}. Codex alone does not grant the Gift. Symbols are optional for parley or omens and do not gate Runekeeper Invocation or the Gift.

\section*{Borrowed Grace}
\label{talent:borrowed-grace}
\index{Talents!Invoker}\index{Imbuement!Lesser}

\textbf{Type:} Invoker Talent — \textit{Lesser Imbuement}

\subsection*{Use}
\begin{itemize}
  \item \textbf{Cost:} 1 Boon or 1 Fatigue, 1 action.
  \item \textbf{Effect (pick one on use):} \textbf{+1 Melee} \emph{or} \textbf{+1 Thematic} (your table’s thematic Skill).
  \item \textbf{Duration:} \textit{Single action/attack} (instantaneous boost).
  \item \textbf{Requirement:} Wield/display the Patron’s \textbf{Symbol}.
  \item \textbf{Obligation:} +1 \textbf{Obligation} to that Patron immediately (see \S\ref{sec:obligation}).
  \item \textbf{Limits:} Cannot be extended, stacked, or \emph{Pushed} for duration.
\end{itemize}

\subsection*{Fictional Framing}
A quick, rule-bending channel through a Patron’s \emph{Symbol}—a sliver of grace, borrowed for a moment and paid for in debt.

\subsection*{Table Guidance (1-liners)}
\begin{itemize}
  \item \textbf{Combat:} Spike a strike vs. a tough foe; or steady a parry in a desperate bind.
  \item \textbf{Skill:} Nudge a pivotal social/ritual/track roll tied to the Patron’s sphere.
  \item \textbf{Fallout:} Repeated use accrues \textbf{Obligation}; NPC faithful may notice “stolen” grace.
\end{itemize}

\subsection*{Balance Notes}
\begin{itemize}
  \item Weaker than full Imbuement: \emph{one} action, no sustain, upfront Obligation.
  \item \textbf{Symbol dependency:} No Symbol, no channel (concealed or lost Symbol = no effect).
\end{itemize}

\subsection*{GM Hooks (quick picks)}
\begin{itemize}
  \item \textbf{Compel Debt:} A Patron agent arrives when Obligation crosses a tick.
  \item \textbf{Clash of Signs:} Using rival Symbols back-to-back risks minor \textbf{Backlash} (drop Position or +1 SB).
  \item \textbf{Spotlight Tell:} Brief visual tell (scent, sigil flare) marks the borrowing to observant NPCs.
\end{itemize}


\centering
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{3.8cm}p{3.8cm}p{7.5cm}@{}}
\toprule
\textbf{Patron} & \textbf{+1 Thematic Skill} & \textbf{Gift / Lore Bestowal} \\
\midrule
Ikasha (Shadow, Penumbra) & Stealth & Grants the hush between footsteps and the raven’s omen at every threshold. \\
Mykkiel (Judgment, Writ) & Command & Grants the authority of seal and sentence, words that bind like iron. \\
The Witness (Truth, Revelation) & Notice & Grants the unblinking gaze that unmasks deceit and remembers every oath. \\
Sealed Gate (Boundaries, Closure) & Tinker & Grants mastery of thresholds—doors that yield or bar at your command. \\
Raéyn (Storm, Tides) & Skirmish & Grants the sailor’s fortune: winds that shift, storms that answer to will. \\
Khemesh (Abyss, Pressure) & Skirmish & Grants the crushing silence of the deep, where strength is drowned in weight. \\
Mab (Glamour, Courts) & Persuade & Grants the mask of favor, a voice that bends courtiers and kindles desire. \\
Sacred Geometry (Perfect Forms) & Tinker & Grants the compass of perfection, every shape reduced to its true measure. \\
Clockwork Monad (Mechanism, Process) & Tinker & Grants the certainty of repetition: a cycle that never falters, a gear that never slips. \\
Varnek Karn (Ossuary, Dominion of the Dead) & Command & Grants the silence of the archive, where the dead obey and records speak. \\
Nidhoggr (Deep Earth, Rot) & Skirmish & Grants the weight of ages, the strength of stone and the hunger of roots. \\
The Traveler (Ways, Roads) & Notice & Grants the open way, a compass that never rests, and roads where none are marked. \\
Oath of Flame \& Light (Dawn, Vows) & Command & Grants the fire of dawn, a vow that shields the faithful and sears the faithless. \\
Carrion King (Carrion, Renewal) & Survival & Grants the feast of decay, where what is dead becomes seed for what lives. \\
Gallows Bell (Doom, Last Rites) & Command & Grants the toll of ending, a voice that closes stories and calls debts due. \\
Old Man of the Black Forest (Primal Humanity, Instinct) & Survival & Grants the wild memory: fang, fire, and the path of instinct through the dark wood. \\
Isoka (Serpents, Shedding) & Skirmish & Grants the serpent’s coil, strength in sudden strike and wisdom in renewal. \\
Inaea (Mercy, Hearth) & Persuade & Grants the hearth’s warmth, shelter to the weary and mercy for the lost. \\
Maelstraeus (Infernal Bargainer) & Persuade & Grants the contract’s weight, every deal sealed in fire and shadow. \\
Livaea (Temptation, Desire) & Persuade & Grants the lure of longing, beauty sharpened into power over hearts. \\
Aliyah, the Chained Angel (Curses, Corruption) & Command & Grants cursed blessings—boons that shine bright yet seed hunger, power that demands a price. \\
\bottomrule
\end{longtable}
\caption{Patron’s Gift: fixed Thematic Skill and lore of their bestowed blessing. Thematic bonuses apply only when the fiction matches the Patron’s domain.}
\label{tab:gift-thematic-map}


\subsection{Specialization vs.\ Mixing}
\label{subsec:mixing}
Characters can mix paths (Summoner, Caster, Invoker, Runekeeper), but specialization is usually stronger and cleaner. Mixing increases upkeep (Obligation, Symbol state, Leash) and action congestion without guaranteed power gains. Let fiction guide choices: Story Beats are prompts to advance the scene, not punishments.

% Patron subsections (split files — keep filenames in the same directory)
\clearpage
\include{malachai.tex}
\clearpage
\include{aveh.tex}
\clearpage
\include{carrion-king.tex}
\clearpage
\include{clockwork-monad.tex}
\clearpage
\include{gallows-bell.tex}
\clearpage
\include{grimmir.tex}
\clearpage
\include{ikasha.tex}
\clearpage
\include{inaea.tex}
\clearpage
\include{index.tex}
\clearpage
\include{inquisitor-prime.tex}
\clearpage
\include{isoka.tex}
\clearpage
\include{khemesh.tex}
\clearpage
\include{livaea.tex}
\clearpage
\include{lunara.tex}
\clearpage
\include{mab.tex}
\clearpage
\include{maelstraeus.tex}
\clearpage
\include{mir-iraath.tex}
\clearpage
\include{morag.tex}
\clearpage
\include{mykkiel.tex}
\clearpage
\include{nidhoggr.tex}
\clearpage
\include{ninth-rim.tex}
\clearpage
\include{oath-flame-light.tex}
\clearpage
\include{pale-shepherd.tex}
\clearpage
\include{palinode.tex}
\clearpage
\include{raeyn.tex}
\clearpage
\include{sacred-geometry.tex}
\clearpage
\include{sealed-gate.tex}
\clearpage
\include{thrysos.tex}
\clearpage
\include{traveler.tex}
\clearpage
\include{varnek-karn.tex}
\clearpage
\include{vorthak.tex}
\clearpage
\include{witness.tex}
\clearpage

\section{Patron Rivalries}
\label{sec:patron-rivalries}

Rivalries set expectations for tone and friction. Use them to color rulings, nudge Position, and guide how Story Beats (SB) land. In their home domains, a Patron’s work tends to start a step better in Position; in a rival’s, a step worse (Keeper’s call).


  \centering
  \renewcommand{\arraystretch}{1.15}
  \begin{longtable}{@{}p{3.4cm}p{3.4cm}p{8.2cm}@{}}
    \toprule
    \textbf{Patron} & \textbf{Primary Rival} & \textbf{Friction in Play (one-line read)} \\
    \midrule
    Raéyn (Sea, Tides, Travel) & Khemesh (Abyssal Maw) & Tides vs. trench: navigation and passage thrive against dread and crushing depths. \\
    Khemesh (Abyssal Maw) & Raéyn (Sea, Tides, Travel) & Abyss unmoors charts: silence, pressure, and alien geometry devour routes. \\
    Sealed Gate (Boundaries, Closure) & The Traveler (Ways, Roads) & Keys vs. roads: jurisdiction and permits against detours and desire lines. \\
    The Traveler (Ways, Roads) & Sealed Gate (Boundaries, Closure) & Paths want to open; gates insist on form—who defines the threshold? \\
    The Witness (Truth, Revelation) & Mab (Glamour, Courts) & Revelation strips glamour; courtly masks fight to endure the gaze. \\
    Mab (Glamour, Courts) & The Witness (Truth, Revelation) & Mask and merriment contest the straight line of testimony. \\
    Ikasha (Shadow, Latent Potential) & The Witness (Truth, Revelation) & Hiding and hush vs. the unblinking eye. \\
    Mykkiel (Judgment, Writ) & Varnek Karn (Necromantic Archives) & Lawful writ and living order against bone-kept precedent and unfinished business. \\
    Varnek Karn (Necromantic Archives) & Oath of Light \& Flame (Dawn, Vows) & Memory of the dead resists purgation by vow and light. \\
    Oath of Light \& Flame (Dawn, Vows) & Khemesh (Abyssal Maw) & Consecrated dawn opposes abyssal hunger and despair. \\
    Sacred Geometry (Order, Pattern) & The Traveler (Ways, Fortune) & Perfect forms vs. opportunistic routes; measure vs. happenstance. \\
    Clockwork Monad (Iteration, Process) & Old Man of the Black Forest (Primal Humanity, Instinct) & Procedure and refinement clash with instinct and feral chaos. \\
    Nidhoggr (Dreaming Antiquity) & Sacred Geometry (Order, Pattern) & Ancient, slumbering memory resists imposed, modern measures. \\
    Carrion King (Carrion, Renewal) & Inaea (Mercy, Hearth) & Rot and renewal against hearth and mercy—life reborn from death or sheltered from it. \\
    Gallows Bell (Doom, Last Rites) & Oath of Light \& Flame (Dawn, Vows) & Doom and execution test the endurance of dawn’s hope and sworn vows. \\
    Old Man of the Black Forest (Primal Humanity, Instinct) & Mab (Glamour, Courts) & The primal face of humanity strips away courtly veneer. \\
    Isoka (Serpents, Shedding) & Sacred Geometry (Order, Pattern) & Serpent coils disrupt the straight line; shedding resists imposed pattern. \\
    Inaea (Mercy, Hearth) & Carrion King (Carrion, Renewal) & Mercy preserves what rot would recycle. \\
    Maelstraeus (Infernal Bargainer) & The Witness (Truth, Revelation) & The bargain lies veiled until truth unmasks it. \\
    Livaea (Temptation, Desire) & Inaea (Mercy, Hearth) & Temptation corrodes hospitality; desire tests sanctuary’s mercy. \\
    Aliyah, the Chained Angel (Curses, Corruption) & Oath of Light \& Flame (Dawn, Vows) & Fallen benedictions twist vows; dawn seeks to burn corruption from blessing. \\
    \bottomrule
  \end{longtable}
  \caption{Primary Patron Rivalries and how they tend to color scenes.}

\include{malachai.tex}
\clearpage
\include{aveh.tex}
\clearpage
\include{carrion-king.tex}
\clearpage
\include{clockwork-monad.tex}
\clearpage
\include{gallows-bell.tex}
\clearpage
\include{grimmir.tex}
\clearpage
\include{ikasha.tex}
\clearpage
\include{inaea.tex}
\clearpage
\include{index.tex}
\clearpage
\include{inquisitor-prime.tex}
\clearpage
\include{isoka.tex}
\clearpage
\include{khemesh.tex}
\clearpage
\include{livaea.tex}
\clearpage
\include{lunara.tex}
\clearpage
\include{mab.tex}
\clearpage
\include{maelstraeus.tex}
\clearpage
\include{mir-iraath.tex}
\clearpage
\include{morag.tex}
\clearpage
\include{mykkiel.tex}
\clearpage
\include{nidhoggr.tex}
\clearpage
\include{ninth-rim.tex}
\clearpage
\include{oath-flame-light.tex}
\clearpage
\include{pale-shepherd.tex}
\clearpage
\include{palinode.tex}
\clearpage
\include{raeyn.tex}
\clearpage
\include{sacred-geometry.tex}
\clearpage
\include{sealed-gate.tex}
\clearpage
\include{thrysos.tex}
\clearpage
\include{traveler.tex}
\clearpage
\include{varnek-karn.tex}
\clearpage
\include{vorthak.tex}
\clearpage
\include{witness.tex}
\clearpage

\subsection{Embracing Corruption}
\index{Corruption!Embracing} \label{subsec:embracing-corruption}

A character with \textbf{2+ levels of Corruption} may choose to \textbf{Embrace} it, transforming creeping decay into a permanent \textbf{Talent}. Embracing never reduces Corruption --- it reshapes it. The deeper the corruption, the greater the power and the cost.

\begin{center}
\begin{longtable}{@{}l l l@{}}
\toprule
\textbf{Corruption Level} & \textbf{Talent Tier} & \textbf{Cost} \\
\midrule
2 (Early) & Minor Talent & +1 Permanent Corruption (cannot Fade below 1) \\
3 (Deepening) & Major Talent & +2 Permanent Corruption, constant visible tell \\
4+ (Severe) & Transcendent Talent & Permanently altered, locked at Stage 3+ \\
\bottomrule
\end{longtable}
\end{center}

\noindent Embraced Talents always reflect the Patron’s themes (e.g., shadow, indulgence, empathy, paranoia). They grant great power, but mark the character irrevocably as claimed.