% Fate's Edge SRD −−− Magic System (Section 10)
% Revised to reflect variable Obligation costs per rite (not flat +1).
% DV formula: DV = max(Obligation Cost − Spirit, Tier)

\section{Magic System}

Magic in Fate's Edge is expressed through four interconnected paths. Each uses the same core resolution engine (Position, DV, Boons, SB) but manages risk differently.

\begin{itemize}
  ℹtem \textbf{Caster (Freeform):} Flexible, improvisational magic. Risk of Backlash.
  ℹtem \textbf{Runekeeper (Rites):} Structured magic through a Patron's Codex. Risk of Obligation.
  ℹtem \textbf{Invoker (Symbols):} Ritual magic through consecrated Symbols. Slow but safe; can be forced.
  ℹtem \textbf{Cantor (Songs):} Perform Low Rites as songs. Risk of Corruption.
\end{itemize}

\subsection{The Four Paths at a Glance}

\begin{tabular}{@{}lcccc@{}}
⊤rule
\textbf{Path} & \textbf{Prerequisite} & \textbf{Roll Skill} & \textbf{Risk Track} & \textbf{Speed} \\
∣rule
Caster & Spellcraft (6 XP) & Arcana & Backlash (1s) & 2 actions (1 for cantrip) \\
Runekeeper & Familiar + Codex & Ritual & Obligation (per Rite) & 1 action \\
Invoker & Patron's Symbol & Ritual & Obligation + Symbol damage & Ritual (slow) or instant (Crack) \\
Cantor & Cantor's Path (8 XP) & Performance + Lore & Corruption (Spirit−based timer) & 1 action (resolves next turn) \\
⊥tomrule
\end{tabular}

\subsection{Caster (Freeform / Weave & Cast)}

Free casters shape raw magic by combining elements and effects. They are flexible but unpredictable.

⫕section{Requirements}
\begin{itemize}
  ℹtem \textbf{Spellcraft (6 XP):} Required for Weave & Cast freeform magic.
\end{itemize}

⫕section{Procedure}
\begin{enumerate}
  ℹtem \textbf{Weave (Action):} Describe the spell and choose its tags (see Section~\ref{sec:tags−effects}).
  ℹtem \textbf{Cast (Action):} Roll Attribute (usually Wits or Spirit) + Arcana. The GM sets DV based on the tags.
\end{enumerate}

\textbf{Cantrip:} A spell with a single tag can be cast as a Cantrip: one roll, standard action, no Weave phase. Cannot bypass wards or create large−scale effects.

⫕section{Spell DV}
\[
\text{DV} = 1 + \text{number of tags}
\]
Dangerous tags (\texttt{[FOLD]}, \texttt{[GATE]}, \texttt{[LEAP]}, \texttt{[CREATE]}, \texttt{[SUMMON]}, \texttt{[TRANSMUTE]}, \texttt{[ANIMATE]}, \texttt{[TRANSFORM]}, \texttt{[DOMINATE]}, \texttt{[REWIND]}) add +2 DV each instead of +1.

⫕section{Backlash}
On a Partial or Miss, or on a Hit that shows two or more 1s, the GM applies backlash colored by the dominant element(s). See Section~\ref{sec:backlash} for details.

\subsection{Runekeeper (Rites)}

Runekeepers bind structured magic through a Codex and a Thiasos (familiar spirit). They specialize in reliable, thematic magic.

⫕section{Requirements}
\begin{itemize}
  ℹtem \textbf{Bond to a single Patron.}
  ℹtem \textbf{Thiasos (Familiar talent, 2 XP):} Enables Patron's Gift.
  ℹtem \textbf{Codex (4 XP):} Grants access to the Patron's Rites and starts the Obligation ledger.
\end{itemize}

⫕section{Invoking a Rite}
\begin{enumerate}
  ℹtem Declare the Rite (1 action).
  ℹtem Mark the Rite's Obligation cost (listed in the rite description, e.g., \texttt{Obligation: 6 segments}).
  ℹtem If the Rite calls for a roll, roll Attribute + Ritual. The GM sets DV using the formula:
  \[
  \text{DV} = \max(\text{Obligation Cost} − \text{Spirit},  \text{Rite Tier})
  \]
  where Rite Tier is 1 for Low, 2 for Standard, 3 for High/Epic.
\end{enumerate}

\textbf{Push It:} Mark +1 additional Obligation to increase duration, range, or potency.

⫕section{Patron's Gift (Imbuement)}
\textit{Requires Thiasos.} Once per scene as 1 action, imbue a held item with +1 Melee and +1 Thematic Skill (determined by your Patron) for the scene. Push It to extend one additional scene (+1 Obligation).

\subsection{Invoker (Symbols)}

Invokers channel power through consecrated Symbols. Rituals are slow but safe; Crack the Seal offers desperate speed.

⫕section{Requirements}
\begin{itemize}
  ℹtem \textbf{One or more Patron's Symbols (4 XP each).} Each Symbol is a Minor Asset.
\end{itemize}

⫕section{Ritual Invocation}
\begin{enumerate}
  ℹtem Display the Symbol (must be visible).
  ℹtem Perform the Rite. Takes \texttt{DV + 1} rounds (Significant Time), where DV is calculated as above.
  ℹtem Mark the Rite's Obligation cost (listed).
  ℹtem If the Symbol is displayed openly during the ritual, reduce the Obligation cost by 1 (minimum 0).
\end{enumerate}

⫕section{Crack the Seal (Instant Cast)}
\begin{enumerate}
  ℹtem Invoke the Rite as 1 action (instead of ritual time).
  ℹtem Mark double the Rite's Obligation cost (or +2 minimum, +3 for High Rites if the base cost is unclear).
  ℹtem The Symbol becomes \textbf{Compromised} (inert). Restore it during downtime (Arcana or Lore test, DV 3) or pay 1 XP.
\end{enumerate}

⫕section{Borrowed Grace (Invoker Talent)}
Once per scene, spend 1 Boon or 1 Fatigue (1 action) to gain +1 die on a combat action, or +1 Effect on a social or stealth action. Requires visibly displaying your Patron's Symbol. If the Symbol is Compromised, Borrowed Grace works at –1 die.

\subsection{Cantor (Songs)}

Cantors perform Low Rites as Songs, channeling power through voice and performance. They track Corruption instead of Obligation.

⫕section{Requirements}
\begin{itemize}
  ℹtem \textbf{Cantor's Path (8 XP).}
  ℹtem \textbf{Performance 2+, Lore 1+, Presence 2+.}
\end{itemize}

⫕section{Singing a Song}
\begin{enumerate}
  ℹtem Begin the Song (1 action). It resolves at the start of your next turn.
  ℹtem Roll Lore + Performance vs DV (typically 2–3).
  ℹtem On success, the Low Rite effect occurs. No Obligation cost.
  ℹtem Mark Corruption accumulation (see below).
\end{enumerate}

⫕section{Push It}
Resolve the Song immediately (instead of next turn). Mark 1 Fatigue and advance your Corruption Timer by 1 segment.

⫕section{Corruption Timer }
Your Corruption Timer has segments equal to your Spirit attribute. Mark a segment when you:
\begin{itemize}
  ℹtem Push a Song.
  ℹtem Perform a Resonant Rite (a powerful thematic working designated by your Patron).
  ℹtem The GM spends a Story Beat on your occult activities.
\end{itemize}

When the timer fills, you immediately gain a thematic boon and burden from the last Patron whose Rite you performed. Then reset the timer to zero (cannot go below cannot go below your current Tier −1 −1).

⫕section{Cantor Corruption Examples (The Eight Core Patrons)}

When a Cantor’s Corruption Timer fills, they immediately gain a thematic \textbf{boon} (benefit) and \textbf{burden} (drawback) from the last Patron whose Rite they performed. The timer resets to zero, but the mark of Corruption remains—a permanent change to the character’s voice, manner, or soul.

The examples below are for the eight patrons recommended for first characters. The GM and player should collaborate to create blooms that fit the fiction.

\begin{longtable}{@{}p{3.2cm}p{4.8cm}p{4.8cm}@{}}
⊤rule
\textbf{Patron} & \textbf{Boon (Benefit)} & \textbf{Burden (Drawback)} \\
∣rule
\endhead
⊥tomrule
\endfoot
\textbf{Oath of Flame & Light} (Dawn & Vows) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Beacon’s Gaze:} Your eyes gleam with dawnlight. Once per session, you may force a single creature to speak truthfully to one question (Resist DV 4).
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Unquenchable:} You cannot knowingly tell a lie. Attempting to do so costs 1 Fatigue and the words burn in your throat.
\end{minipage} \\
∣rule
\textbf{The Traveler} (Ways & Journeys) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Road’s Memory:} You can never be lost on a path you have walked before. +1 die to Navigation rolls on familiar routes.
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Restless Feet:} You cannot stay in one settlement for more than a week without marking 1 Fatigue. The road calls, and you must answer.
\end{minipage} \\
∣rule
\textbf{Inaea} (Angel of the Spider) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Thread−Sight:} You can perceive the emotional bonds between any two people within Near range (hostility, loyalty, love, debt). The GM describes what you see.
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Tangled:} You cannot ignore a plea for help from someone you have previously aided. Refusing costs 1 Fatigue and haunts your dreams.
\end{minipage} \\
∣rule
\textbf{The Witness} (Truth & Revelation) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Unblinking:} You instantly recognize when someone is lying to you. No roll required.
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Cannot Look Away:} You must correct any falsehood you hear, even if doing so is dangerous or socially ruinous. Remaining silent costs 1 Fatigue.
\end{minipage} \\
∣rule
\textbf{Carrion King} (Decay & Renewal) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Rot’s Harvest:} You can sense organic matter in decay within Near range. +1 die to Survival rolls involving scavenging or foraging.
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Compost−Stained:} A faint scent of turned earth follows you. Animals are nervous around you (–1 die to Animal Handling), and you cannot enter consecrated ground without marking 1 Fatigue.
\end{minipage} \\
∣rule
\textbf{Ikasha} (She Who Sleeps) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Shadow’s Daughter:} In dim light or darkness, you may reroll one failed Stealth roll per scene.
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Unseen, Unheard:} Your voice is barely above a whisper. Anyone trying to hear you outside of Close range suffers –1 die to Perception checks involving sound.
\end{minipage} \\
∣rule
\textbf{Grimmir} (The Old Man of the Forest) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Season’s Tongue:} You instinctively know the weather for the next 24 hours and gain +1 die to all Survival rolls in natural terrain.
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Bark−Skinned:} Your skin grows rough and moss−flecked. –1 die to social rolls in civilized settings; you are visibly \textit{wild}.
\end{minipage} \\
∣rule
\textbf{Palinode} (Queen of Encores) &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Perfect Phrase:} Once per session, you may reroll any failed Performance or Sway roll. The second result stands.
\end{minipage} &
\begin{minipage}[t]{4.8cm}
\vspace{0pt}
\textit{Unfinished:} You cannot leave a performance incomplete. If you begin singing, playing, or speaking in public, you must finish or mark 1 Fatigue.
\end{minipage} \\
⊥tomrule
\end{longtable}

\noindent\textbf{Designer’s Note:} These blooms are intended to be \textit{dramatic}, not punishing. A Corruption bloom should feel like a power surge with an interesting complication—not a penalty that discourages play. When a Cantor’s timer fills, celebrate the transformation. The burden is a story hook, not a tax.

⫕section{Inspire Chorus}
Once per scene, while actively singing a Song, all allies within Near range shift Position +1 for one exchange (cannot exceed Controlled).

\subsection{Obligation (Runekeepers and Invokers)}

Obligation measures a Patron's claim on a character. Track it per Patron. The default Obligation capacity is:
\[
\text{Capacity} = \text{Spirit} + \text{Presence}
\]

⫕section{Marking Obligation}
Obligation is marked according to the specific Rite or action:
\begin{itemize}
  ℹtem Invoking a Rite (Runekeeper or Invoker): mark the Rite's listed Obligation cost (e.g., \texttt{Obligation: 6}).
  ℹtem Pushing a Rite: mark +1 additional Obligation (added to the Rite's base cost).
  ℹtem Crack the Seal: mark double the Rite's base cost (or +2 minimum, +3 for High Rites).
  ℹtem Using Rites from two different Patrons in the same scene: mark +1 Obligation to the second Patron.
\end{itemize}

⫕section{Overflow}
If your total Obligation (across all Patrons) exceeds your Capacity:
\begin{itemize}
  ℹtem Mark 1 Fatigue per segment over Capacity.
  ℹtem You cannot invoke new Rites or perform Patron rituals until Obligation is reduced below Capacity.
\end{itemize}

⫕section{Clearing Obligation}
\begin{itemize}
  ℹtem \textbf{Off−screen:} Perform an Act of Service aligned with the Patron (provide an Intricate Description). Clear 1 segment of that Patron's Obligation.
  ℹtem \textbf{On−screen (Focused Devotion talent, 3–5 XP):} Once per scene, spend 1 Boon to clear 1 segment (cannot clear segments accrued in the same scene).
\end{itemize}

\subsection{Backlash (Free Casting)}

Backlash is the unpredictable fallout from shaping raw magic. It occurs on a Partial or Miss, or on a Hit that shows two or more 1s.

⫕section{Backlash Menu (choose one on Partial, two on Miss)}
\begin{itemize}
  ℹtem \textbf{Position Shift:} Your Position worsens by one step for your current or next action.
  ℹtem \textbf{Fleeting Harm/Condition:} Minor injury (1 Fatigue or Harm 1) or a temporary Condition (Dazed, Singed, Disoriented).
  ℹtem \textbf{Exposure/Noise:} Draws unwanted attention — guards alerted, a door slams, a rival notices.
  ℹtem \textbf{Resource Drain:} Lose a component, a tool wears out, or mark 1 Fatigue.
  ℹtem \textbf{Collateral Spark:} A nearby ally, object, or environmental feature is threatened (GM describes).
\end{itemize}

All backlash is colored by the dominant element(s) used in the spell. See the Elemental Backlash table in Section~\ref{sec:tags−effects}.

\subsection{Obligation Capacity and Overflow Summary}

\[
\text{Capacity} = \text{Spirit} + \text{Presence}
\]

\begin{itemize}
  ℹtem \textbf{Exceeding Capacity:} Mark 1 Fatigue per segment over. Cannot invoke new Rites or perform Patron rituals until reduced.
  ℹtem \textbf{Clearing:} Act of Service (off−screen, 1 segment) or Focused Devotion talent (on−screen, 1 Boon per segment).
\end{itemize}

\subsection{Corruption Timer Summary (Cantor)}

\[
\text{Timer Size} = \text{Spirit}
\]

\begin{itemize}
  ℹtem Mark a segment when you Push a Song, perform a Resonant Rite, or when the GM spends SB on your occult activities.
  ℹtem When full: gain a thematic boon and burden from your last Patron; reset timer to zero.
\end{itemize}