% --- Fate's Edge SRD — Section 10: Tags & Effects Library ---
\section{Tags \& Effects Library}
\label{sec:tags}

Tags are keywords that define the mechanical and narrative behavior of spells, rites, talents, and assets. 
They do nothing on their own; they only function when granted by a rule (Spell, Rite, Talent, Ability). 
Unless noted otherwise, the **Duration** of a tag effect is **Scene**, and the **DV** for contests or resistances is **set by the fiction** (typical range 2–5).

\subsection{How to Read a Tag Entry}
Each entry includes:
\begin{itemize}
  \item \textbf{Name \& Effect:} What the tag does.
  \item \textbf{DV Guideline:} Typical Difficulty Value when the tag is contested.
  \item \textbf{Outcomes (Hit / Partial / Miss):} Standard resolution guidance.
  \item \textbf{Countered by:} Common ways to negate or bypass the effect.
\end{itemize}

\subsection{Control \& Countermagic}

\paragraph{[DISPEL] – End an ongoing magical effect or construct.}
\begin{itemize}
  \item \textbf{DV:} By fiction (usually equal to the effect’s Tier or casting DV).
  \item \textbf{Hit:} Effect ends or is suppressed per source text.
  \item \textbf{Partial:} Effect suppressed briefly (one exchange) or reduced in scope.
  \item \textbf{Miss:} No effect; GM may gain SB.
\end{itemize}

\paragraph{[COUNTER] – Interrupt a spell, rite, or ritual in progress.}
\begin{itemize}
  \item \textbf{Timing:} During the casting window (declared action or ritual phase).
  \item \textbf{DV:} By fiction (complexity of the interrupted effect).
  \item \textbf{Hit:} Cancel the action; resources used are lost.
  \item \textbf{Partial:} Worsen the caster’s Position or impose Fatigue.
  \item \textbf{Miss:} No interruption; GM may spend SB to escalate.
\end{itemize}

\paragraph{[BARRIER] – Create cover, obstruction, or a physical wall.}
\begin{itemize}
  \item \textbf{DV:} By fiction (size, material, duration).
  \item \textbf{Hit:} Place a barrier with Integrity Clock [2/4/6]; it blocks movement/attacks until cracked.
  \item \textbf{Partial:} Barrier is fragile or has a gap.
  \item \textbf{Countered by:} [DISPEL], brute force (Athletics), or clever route.
\end{itemize}

\paragraph{[SEAL] / [UNSEAL] – Lock or unlock a container, door, or portal (not spirits).}
\begin{itemize}
  \item \textbf{DV:} By fiction (quality of lock, magical binding).
  \item \textbf{Hit:} Locked/unlocked as intended; specify any bypass conditions.
  \item \textbf{Partial:} Seal is temporary or leaks (can be forced open with +1 DV).
\end{itemize}

\subsection{Concealment \& Revelation}

\paragraph{[VEIL] – Obscure a person, object, or zone.}
\begin{itemize}
  \item \textbf{DV:} By fiction (scale, conditions).
  \item \textbf{Hit:} Impose –1 die to Notice or perception rolls against the veiled subject; specify limits (angle, distance, scent).
  \item \textbf{Partial:} Works only at a limited range or in dim light.
  \item \textbf{Countered by:} [REVEAL], strong light, physical proof.
\end{itemize}

\paragraph{[REVEAL] – Expose illusions, disguises, or hidden clauses.}
\begin{itemize}
  \item \textbf{DV:} By fiction (complexity of the concealment).
  \item \textbf{Hit:} Surface the truth; describe how the revelation manifests (brief glow, visible flaw, etc.).
  \item \textbf{Partial:} A tell or clue, not the full picture.
\end{itemize}

\paragraph{[MARK] – Tag a target for tracking or leverage.}
\begin{itemize}
  \item \textbf{DV:} By fiction (resistance, visibility).
  \item \textbf{Hit:} Place a visible or invisible mark; once per scene, you or an ally gain +1 die when acting directly against the Marked target.
  \item \textbf{Partial:} Mark is noisy or short‑lived.
  \item \textbf{Countered by:} [CLEANSE], [DISPEL], or listed materials (salt, iron).
\end{itemize}

\subsection{Boons \& Burdens}

\paragraph{[CURSE] – Inflict a sticky hindrance with a clear release condition.}
\begin{itemize}
  \item \textbf{DV:} By fiction (severity, target’s resilience).
  \item \textbf{Hit:} Apply a named Condition (e.g., \textsc{Shaken}, \textsc{Weakened}) that persists until the release condition is met.
  \item \textbf{Partial:} Condition is milder or intermittent.
  \item \textbf{Countered by:} [CLEANSE] or the listed key (apology, theft return, pilgrimage).
\end{itemize}

\paragraph{[CLEANSE] – Remove or suppress a Condition (poison, disease, curse, fear).}
\begin{itemize}
  \item \textbf{DV:} By fiction (potency of the affliction).
  \item \textbf{Hit:} Remove one named Condition.
  \item \textbf{Partial:} Suppress or reduce severity for the scene.
\end{itemize}

\paragraph{[FORTIFY] – Harden a person, place, or object against a specific vector (fire, blades, fear, persuasion).}
\begin{itemize}
  \item \textbf{DV:} By fiction (scale, strength).
  \item \textbf{Hit:} Raise Position by one step or reduce consequence severity for attacks of that vector this scene.
  \item \textbf{Partial:} Limited scope (one ally, one doorway).
\end{itemize}

\subsection{Influence \& Oaths}

\paragraph{[COMMAND] – Issue a clear order to a sapient target.}
\begin{itemize}
  \item \textbf{DV:} By fiction (authority, leverage, fear, ritual standing).
  \item \textbf{Hit:} Target complies now or suffers an immediate cost stated by the GM.
  \item \textbf{Partial:} Target hesitates, bargains, or partially complies.
  \item \textbf{Miss:} Refusal or blowback (GM may spend SB).
\end{itemize}

\paragraph{[OATH] – Bind parties to terms; breaking has teeth.}
\begin{itemize}
  \item \textbf{DV:} By fiction (witnesses, sanctity, stakes).
  \item \textbf{Hit:} Enforceable pact with stated boon for keeping and consequence for breaking.
  \item \textbf{Partial:} A loophole exists in the oath.
  \item \textbf{Countered by:} Release clause, [CLEANSE] if text allows.
\end{itemize}

\paragraph{[SANCTIFY] – Consecrate a zone to a code or Patron.}
\begin{itemize}
  \item \textbf{DV:} By fiction (size, opposition).
  \item \textbf{Hit:} Specify allowed/prohibited acts; violations start at Controlled Position or impose a soft consequence.
  \item \textbf{Partial:} Patchy coverage (e.g., only the altar, not the whole room).
  \item \textbf{Countered by:} Opposing rite, profanation (bloodshed, lies), time.
\end{itemize}

\subsection{Movement \& Making}

\paragraph{[PASSAGE] – Declare a route as permitted or easy.}
\begin{itemize}
  \item \textbf{DV:} By fiction (length, danger).
  \item \textbf{Hit:} Allies on that path gain improved flow (+1 Position or ignore one level of difficult terrain).
  \item \textbf{Partial:} Only one ally or one segment benefits.
\end{itemize}

\paragraph{[TRANSPORT] – Move a target across an obstacle (blink, lift, pull).}
\begin{itemize}
  \item \textbf{DV:} By fiction (mass, range, warding).
  \item \textbf{Hit:} Relocate within listed range; state any tells or costs (SB, Fatigue).
  \item \textbf{Partial:} Arrive off‑balance, drop gear, or suffer worse Position.
\end{itemize}

\paragraph{[CONJURE] – Create a useful object, cover, or hazard.}
\begin{itemize}
  \item \textbf{DV:} By fiction (size, complexity).
  \item \textbf{Hit:} Conjure an item or zone with Integrity Clock [2/4/6] (if destructible) or a ticking hazard (burn, freeze, haze).
  \item \textbf{Partial:} Fragile or short‑lived (lasts one exchange).
  \item \textbf{Countered by:} [DISPEL], force, time.
\end{itemize}

\subsection{Outsider‑Specific Tags}
\label{sec:tags-outsiders}

\paragraph{[WARD] – Challenge Outsiders crossing a warded edge or zone.}
\begin{itemize}
  \item \textbf{DV = target Cap.} (Cap is the Outsider’s tier: 1 for Lesser, 3 for Greater.)
  \item \textbf{Hit:} Crosses; add +DV segments to its Leash.
  \item \textbf{Partial:} Crosses; add +1 segment.
  \item \textbf{Miss:} Cannot cross this beat.
\end{itemize}

\paragraph{[BANISH] – Drive a visible Outsider toward departure.}
\begin{itemize}
  \item \textbf{DV = target Cap.}
  \item \textbf{Hit:} Add +DV to its Leash.
  \item \textbf{Partial:} Add +1.
  \item \textbf{Miss:} No effect.
\end{itemize}

\paragraph{[UNWARD] – Unmake or suppress a [WARD] created by a Talent, Ability, or Spell.}
\begin{itemize}
  \item \textbf{DV:} By fiction (strength of the ward).
  \item \textbf{Hit:} Dismiss or suppress per source text.
  \item \textbf{Partial:} Weaken or suppress briefly (one exchange).
  \item \textbf{Miss:} No effect.
\end{itemize}

\subsection{Elemental Backlash Coloring (Quick Reference)}
When a magical roll generates SB (1s) or when a Partial/Miss occurs, the GM may apply a thematic backlash based on the dominant element(s). Use this table for inspiration:

\centering
\begin{longtable}{l l}
\toprule
\textbf{Element} & \textbf{Minor / Major Backlash Examples} \\
\midrule
Earth & Rubble, pin, heavy footing; vs Air: sound carries, exposure \\
Fire & Burns, flares; vs Water: slick, sputter, dim \\
Air & Scatter, drop, vertigo; vs Earth: stuck, dust choke \\
Water & Leak, flood, cold drag; vs Fire: smoke, shorted gear \\
Fate & Options close, “only one way”; vs Luck: mischance hits ally \\
Luck & Side‑effect elsewhere, fragile success; vs Fate: harsher fixed outcome \\
Life & Overgrowth, fever; vs Death/Dreams: numbness, sleep‑tug \\
Death/Dreams & Fade, threshold opens/closes; vs Life: pain returns, rot \\
\bottomrule
\end{longtable}

\subsection{Stacking \& Duration Rules (Global)}
\begin{itemize}
  \item \textbf{No Same‑Source Stacking:} Multiple scene‑long effects from the same source do not stack; take the best instance.
  \item \textbf{Identical Tag Collision:} If two abilities would grant the same Tag in the same window, use the stronger; the weaker is suppressed while it lasts.
  \item \textbf{Default Duration:} Scene, unless a source states otherwise (e.g., “until the end of your next turn”).
  \item \textbf{DV by Fiction:} Potency, preparation, place, materials, and opposition set DV (typical 2–4; 5+ for set‑pieces).
\end{itemize}