
% --- Fate's Edge SRD — Section 11: Backlash, Obligation, and Ritual Consequences ---
% Include this file from your main .tex with: \input{11-backlash-obligation.tex}

\section{Backlash, Obligation, and Ritual Consequences}
\label{sec:backlash-obligation}

This section codifies how magical risks bite and how Patron debts accrue. It unifies the fallout for \textbf{Casting (freeform)}, \textbf{Rites (warlock path)}, and \textbf{Invocations (symbol path)} in a way that is fully consistent with the core SB/Boon economy.

\subsection{Backlash (Casting: Weave \& Cast)}
Backlash colors the cost of magic and is always expressed through fiction first.
\begin{itemize}
  \item \textbf{Trigger Windows:}
    \begin{itemize}
      \item \emph{Primary:} On a \textbf{Partial/Miss} on either the \emph{Weave} or \emph{Cast} roll.
      \item \emph{Also:} On a \emph{Hit} that shows \textbf{two or more 1s}, minor backlash may ride alongside success.
    \end{itemize}
  \item \textbf{Expression:} Choose a consequence flavored by the spell's Element (see \S\ref{sec:tags}, Elemental Backlash). Options include: worsen Position by 1; impose a condition or brief Harm; drain time/resources; alert foes; create a fragile side-effect.
  \item \textbf{SB Integration:} Backlash does \emph{not} generate extra SB by itself; it is one way the GM spends the SB created by rolled 1s (see \S2). The color (Fire, Fate, etc.) guides \emph{how} that SB spend lands.
  \item \textbf{Scale:} Small/local effects (DV 2) prefer small backlash; bigger spans (DV 4--5+) invite larger, more persistent fallout.
\end{itemize}

\paragraph{Backlash Menu (guideline)}
Pick one on \emph{Partial}; pick two on \emph{Miss}. Always color by Element.
\begin{itemize}
  \item \textbf{Position Shift} ($-1$ step) for your current or next action.
  \item \textbf{Fleeting Harm/Condition} (e.g., sear, vertigo, chill) that matters for this scene.
  \item \textbf{Exposure/Noise} draws notice or complicates stealth.
  \item \textbf{Resource Drain} (time, focus, a component damaged).
  \item \textbf{Collateral Spark} threatens an ally or fragile thing nearby.
\end{itemize}

\subsection{Obligation (Rites: Warlock Path)}
Rites bind you to a Patron through \textbf{Obligation}. Each Patron has its own clock.
\begin{itemize}
  \item \textbf{Marking Segments:} Invoking a Rite marks its listed Obligation (typically $+1$ segment). \emph{Push It} adds $+1$ more.
  \item \textbf{Cross-Patron Interference:} If you Invoke Rites from two different Patrons in the same scene, immediately mark $+1$ Obligation to the \emph{second} Patron (jealous attention).
  \item \textbf{Resolution:} When a clock fills, the GM resolves debt in-fiction (audits, omens, service, seizure of a boon). Reduce/clear by fitting service or downtime actions.
\end{itemize}

\paragraph{Patron’s Gift (Imbuements) and Obligation}
\begin{itemize}
  \item \textbf{Activation:} 1 Action once/scene to empower an item with +1 Weapon (Melee) and +1 Thematic Skill for the scene.
  \item \textbf{Push It:} Extend for \emph{one} additional scene; mark $+1$ Obligation to that Patron.
\end{itemize}

\subsection{Invocation via Symbols (Invoker Path)}
Symbols allow ritual access to a Patron's Rites without a bond.
\begin{itemize}
  \item \textbf{Access:} Each \textbf{Patron’s Symbol (Minor Asset, 4 XP)} grants ritual access to that Patron's Rite list.
  \item \textbf{Ritual Completion:} Completing an Invoker ritual \emph{always} marks $+1$ Obligation on that Rite's clock, even outside scene context (attention cost).
  \item \textbf{No Push:} Invoker rituals cannot use \emph{Push It}.
  \item \textbf{Environment:} Requires a suitable ritual space; the Symbol must be openly displayed during the rite.
\end{itemize}

\paragraph{Crack the Seal (Instant Cast)}
Convert a ritual into a one-action cast by compromising the Symbol.
\begin{itemize}
  \item \textbf{Cost:} Set the Symbol to \emph{Compromised} and mark $+2$ segments on that Rite's Obligation clock (\emph{$+3$ if High-Power}).
  \item \textbf{Instability:} The GM may immediately spend \textbf{1 SB} on-theme to reflect ritual instability.
  \item \textbf{Aftermath:} Restore during downtime with a fitting test (DV 3 or by fiction). \emph{Success:} Maintained. \emph{Shaky:} Neglected (future rituals add $+1$ Obligation until properly restored).
\end{itemize}

\paragraph{Rival Symbols and Mixed Use}
\begin{itemize}
  \item \textbf{Rival Symbol Penalty:} Displaying or carrying a rival Patron's Symbol while Invoking another Patron’s Rite worsens Position by 1 for that action and immediately marks $+1$ Obligation to the current Patron; on any rolled \textbf{1}, the GM may also trigger a Patron omen/glitch.
  \item \textbf{Multi-Symbol Overload:} Carrying \textbf{4+} different Symbols causes choir effects: the first Invoker ritual completed each scene marks $+1$ additional Obligation.
  \item \textbf{Concurrency:} An Invoker may sustain at most \textbf{Spirit} ongoing Invoker rituals. Starting a new one either ends the oldest or marks $+1$ Obligation on it.
\end{itemize}

\subsection{Interruption \& Counterplay}
\begin{itemize}
  \item \textbf{[COUNTER]} can interrupt \emph{Weave}, \emph{Cast}, or a \emph{Rite} within its casting window (DV by fiction).
  \item \textbf{[DISPEL]} ends or suppresses ongoing effects (DV by fiction).
  \item \textbf{[UNWARD]} suppresses/dismisses [WARD]s; \textbf{[BANISH]} and [WARD] interact with Outsiders per \S\ref{sec:tags-outsiders}.
\end{itemize}

\subsection{GM Budget Dials (Practical Limits)}
Keep pressure high but readable.
\begin{itemize}
  \item \textbf{SB Budgets:} Use scene/session limits in \S7.
  \item \textbf{Obligation Pace:} A typical session sees 2--4 segments marked for an active warlock; spikes occur on set-pieces or heavy \emph{Push} use.
  \item \textbf{Backlash Visibility:} Favor big, legible consequences over many minor pinpricks.
\end{itemize}

\subsection{Micro-Examples}
\paragraph{Fire Cast, Partial}
You \emph{Weave} flame to blind a squad (DV 3). Partial with two 1s. GM spends SB to \emph{Position -1} (flare blinds you too) and colors backlash as singed lashes; patrol is alerted (\emph{Exposure}).

\paragraph{Warlock Push and Debt}
You Invoke \emph{Circle of Denial [WARD]} and \emph{Push It} to harden the ring. Mark $+1$ Obligation for the Rite plus $+1$ for the push. When a demon tests the ring, use [WARD] vs Cap; on its Hit, add $+DV$ to its Leash.

\paragraph{Crack the Seal Under Fire}
You present Ikasha’s Symbol and \emph{Crack the Seal} to lay an instant shadow lane. Symbol $\rightarrow$ Compromised; mark $+2$ Obligation. GM immediately spends 1 SB to dim all lights (panic), then the lane forms. During downtime, you restore the Symbol (Arcana DV 3): a shaky hit leaves it \emph{Neglected} until you perform the full rite of cleaning.
