% --- Fate's Edge SRD — Section 12: Character Creation & Starting Build ---
% Include this file from your main .tex with: \input{12-character-creation.tex}

\section{Character Creation \& Starting Build}

\subsection{Starting Build Points}
Players begin with \textbf{30 XP} to allocate during initial character creation. This represents a balanced baseline for competent starting characters.

\subsection{Enhanced Starting Builds}
Players may exceed the standard 30 XP build through narrative engagement:
\begin{itemize}
  \item \textbf{Bonds:} Up to two player-defined mutual bonds may be taken for \textbf{+2 XP total}. (See bonds and bond-driven play in the core rules.)
  \item \textbf{Complications:} Up to two initial complications may be accepted for \textbf{+4 XP total}. \emph{Note:} Scenes start with \textbf{+1 banked SB per complication per character} until those complications have cleared.
\end{itemize}
This allows for a maximum starting build of \textbf{34 XP}. Players are encouraged to aim for 30 XP and use bonds/complications to mitigate slight overages while maintaining narrative balance.

\subsection{Recommended Approach}
GMs should encourage players to:
\begin{itemize}
  \item Target 30 XP for balanced starting characters.
  \item Use bonds and complications to enhance characterization rather than pure mechanical optimization.
  \item Consider the narrative implications of any starting advantages.
\end{itemize}

\subsection{Initial Complications (Reminder)}
For each \emph{initial complication} taken at character creation:
\begin{itemize}
  \item Start each scene with \textbf{+1 banked SB} \emph{per character with initial complications} until those complications have cleared in play.
\end{itemize}

\subsection{Early Talent Choices (Magic Access)}
\begin{itemize}
  \item \textbf{Caster’s Gift (2 XP):} Required for Weave \& Cast freeform magic. Provides flexible but limited effects with Element-colored backlash.
  \item \textbf{Familiar / Thiasos (2 XP):} Required for Patron features such as \emph{Patron’s Gift} (Imbuements).
  \item \textbf{Codex (4 XP):} Required to fully join a Patron’s service as a Runekeeper (Rites access, Obligation). A Runekeeper may serve only one Patron.
  \item \textbf{Patron’s Symbol (4 XP):} Minor Asset. Grants ritual access to a Patron’s Rites via the Invoker path. One Symbol per Patron; Invokers may carry Symbols of multiple Patrons.
\end{itemize}

\subsection{Quick Build Guides}
\paragraph{Caster (Freeform)}
Buy \emph{Caster’s Gift (2 XP)}. Invest in Arcana-adjacent attributes/skills. Expect backlash; leverage aligned loci, tools, and prep for DV/Effect nudges. Casting is flexible but less reliable than Rites.

\paragraph{Runekeeper (Rites User)}
Acquire \emph{Thiasos (Familiar)} and \emph{Codex (4 XP)} tied to a single Patron. You can Push Rites for $+1$ Obligation. Use \emph{Patron’s Gift (Imbuement)} once per scene (activation: 1 action). The Gift provides exactly \textbf{+1 Melee} and \textbf{+1 Thematic Skill}, fixed by Patron (see Patron table).

\paragraph{Invoker (Symbol Path)}
An Invoker purchases one or more \emph{Patron’s Symbols} (4 XP each). Each Symbol grants ritual access to that Patron’s Rites. 
\begin{itemize}
  \item \textbf{Ritual Invocation.} Performing a Rite via Symbol takes $\text{DV} + 1$ rounds and always marks $+1$ Obligation. 
  \item \textbf{Crack the Seal.} Cast instantly by setting the Symbol to \textsc{Compromised} and marking $+2$ Obligation ($+3$ if High-Power). 
  \item \textbf{Multiple Symbols.} Invokers may hold Symbols from multiple Patrons; each Symbol tracks its own state (Normal, Wary, Compromised, Shattered). 
\end{itemize}

\paragraph{Summoner (Pact-Whisperer)}
You call spirits fast and manage their timer. \emph{Call} a spirit (1 action) $\rightarrow$ \emph{Bind} it by spending \textbf{1 Boon} or marking \textbf{1 Fatigue} $\rightarrow$ track its \textbf{Leash = Cap + Command} segments. The Leash ticks when the spirit takes harm, you split focus (you act meaningfully while it acts), you order it against its nature, rivals contest it, or it rushes \emph{Close} to \emph{Far}. Once per round you may spend \textbf{1 Boon} to clear \textbf{1} Leash tick (Boon Finesse). Limit \textbf{one active spirit} at a time; issuing a meaningful command uses your action. Spirits depart at downtime unless an ability states otherwise.

\subsection{Session Zero Tips}
\begin{itemize}
  \item Establish the table’s tone and stakes; tie Drives to setting fronts.
  \item Map Bonds among PCs; mark possible bond-driven Boon triggers.
  \item Seed 1--2 personal Complications per PC for early spotlight.
  \item Discuss Patron choices, Gifts, and what their omens look like.
\end{itemize}