% Fate's Edge SRD −−− Rites & Invokers (Section 12)

\section{Rites and Invokers}

This chapter covers universal rituals (available to any character with time and components) and the Invoker path (characters who channel Patrons through consecrated Symbols).

\subsection{Universal Rituals}

Any character, regardless of magical talent, may attempt a universal ritual. Requirements:

\begin{itemize}
  ℹtem A clear intent (what you want to achieve).
  ℹtem A method (e.g., summoning circle, prayer, sympathetic object).
  ℹtem Components (consumed or used up).
  ℹtem Time (at least one scene of uninterrupted work).
\end{itemize}

⫕section{Ritual Resolution}
The GM sets a DV (typically 4–6) based on scope. The player rolls \textbf{Spirit + Lore} (or another appropriate skill).

\begin{itemize}
  ℹtem \textbf{Clean Success:} The ritual works as intended.
  ℹtem \textbf{Success & Cost:} Works, but with a cost (backlash, extra time, compromised component). GM spends SB.
  ℹtem \textbf{Partial:} Partial effect; player gains 1 Boon.
  ℹtem \textbf{Miss:} Fails; GM may impose a complication (wrong spirit answers, ward backfires). Player gains 2 Boons.
\end{itemize}

⫕section{Universal Rituals Table}

\begin{tabular}{@{}lllp{4cm}@{}}
⊤rule
\textbf{Ritual} & \textbf{Tier} & \textbf{Cast Time} & \textbf{Components & Effect} \\
∣rule
Wayfinder's Thread & Low & 1 min & Red cord knotted thrice; whisper destination. Create tether to safe path. Cost: Fatigue if used $>$1/scene. Push: +1 SB reveals shortcut. \\
Oath−Ward & Low & 5 min & Chalk circle; sworn phrase. Ward small area vs intrusion. Cost: Breaking oath triggers SB. \\
Ember−Glass & Low & 1 min & Smoked glass. Sense heat sources through cover. Cost: lose torch/tinder. \\
Salt−Cut & Low & 1 min & Salt, bronze knife. Sever simple ongoing effect (rope, charm). Cost: consume salt. \\
River's Memory & Medium & 10 min & Bowl, personal token. Scry recent event tied to token (blurry). Cost: token ruined. \\
Bargain−Bead & Medium & 10 min & Two carved beads. Invite spirit to parley. Cost: give valuable concession or take SB. \\
Quiet Veil & Medium & 5 min & Ash, muffled bell. Muffle group sounds for scene. Cost: Muted Condition. \\
Dream−Way Marker & Medium & 10 min & Inked sleep mask, water cadence. Mark safe entrance to Ways Between; short transit. Cost: all Shaken on waking. \\
Purge & Medium & 10 min & Bitter herbs, clean blade. Cleanse taint/curse one step. Cost: caster takes minor echo (–1 next action). \\
Summoner's Gate & High & 20 min & Circle, true−name sigil. Call known entity safely. Cost: occupies concurrency slot; breaking terms creates SB. \\
⊥tomrule
\end{tabular}

\subsection{Invokers (Symbol Path)}

Invokers channel power through consecrated \textbf{Symbols} — physical tokens tied to a specific Patron.

⫕section{Requirements}
\begin{itemize}
  ℹtem \textbf{Patron's Symbol (4 XP each):} Minor Asset. Grants ritual access to that Patron's Rites.
  ℹtem One Symbol per Patron; Invokers may carry Symbols of multiple Patrons.
\end{itemize}

⫕section{Ritual Invocation}
\begin{itemize}
  ℹtem Takes \textbf{DV + 1 rounds} to complete.
  ℹtem Display the Symbol openly (visible).
  ℹtem Always marks \textbf{+1 Obligation} to that Patron.
  ℹtem If the Symbol is displayed, DV –1 and Obligation –1 (minimum 0 Obligation).
\end{itemize}

⫕section{Crack the Seal (Instant Cast)}
Resolve the Rite immediately (1 action) by:
\begin{itemize}
  ℹtem Marking \textbf{+2 Obligation} (+3 for High Rites).
  ℹtem Setting the Symbol to \textbf{Compromised} (inert until restored).
  ℹtem The GM may immediately spend 1 SB on a thematic complication.
\end{itemize}

⫕section{Restoring a Symbol}
\begin{itemize}
  ℹtem Requires a Downtime action.
  ℹtem Roll \textbf{Wits + Arcana or Lore} (DV 3).
  ℹtem \textbf{Clean Success:} Symbol returns to Maintained.
  ℹtem \textbf{Partial:} Symbol becomes Neglected (future rituals +1 Obligation).
  ℹtem \textbf{Miss:} Symbol remains Compromised; requires a quest or Patron service to restore.
\end{itemize}

⫕section{Multiple Symbols}
\begin{itemize}
  ℹtem Carrying 4 or more different Symbols causes the first Invoker ritual each scene to mark +1 additional Obligation.
  ℹtem Displaying a rival Patron's Symbol while invoking another worsens Position by 1 and marks +1 Obligation to the current Patron. On any rolled 1, the GM may trigger a Patron omen or glitch.
\end{itemize}

\subsection{Borrowed Grace (Invoker Talent)}
\textbf{Cost:} 3–5 XP (minor talent). Once per scene, spend 1 Boon or 1 Fatigue (1 action) to gain:
\begin{itemize}
  ℹtem +1 die to a single combat action, \textbf{or}
  ℹtem +1 Effect to a social or stealth action.
\end{itemize}
Requires visibly displaying your Patron's Symbol. If the Symbol is Compromised, Borrowed Grace works at –1 die.

\subsection{The Traveler (Ways & Journeys)}

The Traveler grants rites of safe passage, wayfinding, and road covenants. All rites below are available to Runekeepers and Invokers sworn to The Traveler.

⫕section{Low Rites}

\begin{description}
  ℹtem[Rite of Road−Sense (Low, Obligation: 4)] ℎfill \\
  \textit{Instant; Self; No.} \\
  \textbf{Tags:} \texttt{[NAVIGATION]}, \texttt{[AWARENESS]} \\
  \textbf{Effect:} You unerringly pick the fastest safe route in Near/Far range. Gain +1 die to avoid ambushes or delays. Create a 4−segment \textit{Path Memory} timer to ignore one instance of difficult terrain. \\
  \textbf{Push It:} Spot one hidden bypass, but generate 1 SB (Clubs). \\
  \textbf{DV:} \texttt{max(4 – Spirit, 1)}.

  ℹtem[Rite of Traveler’s Boon (Low, Obligation: 5)] ℎfill \\
  \textit{Scene; Self/Ally; No.} \\
  \textbf{Tags:} \texttt{[MOVEMENT]}, \texttt{[BOND]}, \texttt{[HASTE]} \\
  \textbf{Effect:} Ignore one level of difficult terrain or bureaucratic delay. +1 Effect to travel or escape checks. If shared, create a 2−segment \textit{Shared Journey} bond with an ally. \\
  \textbf{Push It:} Extend to one more ally; mark 1 SB (Diamonds). \\
  \textbf{DV:} \texttt{max(5 – Spirit, 1)}.
\end{description}

⫕section{Standard Rite}

\begin{description}
  ℹtem[Rite of the Waymark (Standard, Obligation: 8)] ℎfill \\
  \textit{Scene; Near; No.} \\
  \textbf{Tags:} \texttt{[MARK]}, \texttt{[GUIDANCE]}, \texttt{[PASSAGE]} \\
  \textbf{Effect:} Declare a lane as permitted or easy. Allies gain improved Position/Effect or ignore one obstacle. Create a 6−segment \textit{Marked Path} timer. \\
  \textbf{Push It:} The lane persists between scenes; the first enemy to exploit it forces 1 SB (Spades). \\
  \textbf{DV:} \texttt{max(8 – Spirit, 2)}.
\end{description}

\noindent \textbf{Note:} For all The Traveler’s rites, the Difficulty Value is calculated as \texttt{DV = max(Obligation Cost – Spirit, Rite Tier)} where Rite Tier is 1 for Low, 2 for Standard. Spirit reduces both the roll difficulty and the risk of Obligation overflow.

\subsection{Additional Sample Rites by Patron}

These Low and Standard Rites are examples. See the full Player's Guide for complete lists.

⫕section{Oath of Flame & Light (Dawn, Vows)}
\begin{itemize}
  ℹtem \textbf{Rite of Kindle Vow (Low):} Bind a scene−long vow; bearer gains +1 die when fulfilling it.
  ℹtem \textbf{Rite of Lay on Hands (Low):} Cleanse affliction, reduce Harm by 1, or remove 1 Fatigue.
  ℹtem \textbf{Rite of Radiant Smite (Standard):} Your next melee strike flares with dawnfire. Upgrade Effect by one step, add +1 Burn Harm, or −1 die to their next roll. Against undead, oath‑breakers, or outsiders: oath‑breakers suffer +1 Harm \textit{and} –1 die, outsiders gain +1 leash ticks. \\
  .
\end{itemize}

⫕section{The Sealed Gate (Thresholds, Wards)}
\begin{itemize}
  ℹtem \textbf{Rite of Sealed Threshold (Low):} Mark a doorway; unauthorized crossing requires Resolve test (DV 3).
  ℹtem \textbf{Rite of Circle of Denial (Standard):} Protective circle; spirits/outsiders must resist (Resolve vs DV 4) to enter.
  ℹtem \textbf{Rite of Banishment Knot (High):} Cast out a supernatural entity; test vs its Cap.
\end{itemize}

⫕section{Ra\'eyn, Mistress of the Sea (Tides, Storms)}
\begin{itemize}
  ℹtem \textbf{Rite of Tidemark's Blessing (Low):} Steady footing on wet surfaces; +1 die to balance/climb.
  ℹtem \textbf{Rite of Whispering Currents (Low):} Learn safest route on water; +1 die to navigation.
  ℹtem \textbf{Rite of Wave−Road Blessing (Standard):} Consecrate a sea lane; +2 dice to travel or evasion.
\end{itemize}

⫕section{Ikasha, She Who Sleeps (Shadow, Secrets)}
\begin{itemize}
  ℹtem \textbf{Touch the Umbral Veil (Low):} +1 die to one Stealth roll; soften a noise.
  ℹtem \textbf{Rite of Crossroads Raven (Low):} Omen−raven gives +1 die to navigation or forces enemy hesitation.
  ℹtem \textbf{Become the Shadow at the Crossroads (High):} Intangible to mundane harm; pass through gaps; +2 dice Stealth.
\end{itemize}

\subsection{Obligation and Invokers}

Invokers track Obligation per Patron (see Section~\ref{sec:backlash−obligation}). Each ritual invocation marks +1 Obligation (reduced by displaying the Symbol). Crack the Seal marks +2 Obligation (+3 for High Rites). Obligation Capacity = \textbf{Spirit + Presence}. Exceeding capacity causes Fatigue.

\subsection{Ritual Interruption and Counterplay}
\begin{itemize}
  ℹtem A ritual can be interrupted by damage, loud noise, or direct countermagic before completion.
  ℹtem A character may use \texttt{[COUNTER]} (see Section~\ref{sec:tags−effects}) to disrupt a ritual in progress.
  ℹtem If interrupted, the ritual fails; components may be lost. The GM gains 1 SB.
\end{itemize}

