
% --- Fate's Edge SRD — Section 13: Advancement & XP ---
% Include this file from your main .tex with: \input{13-advancement-xp.tex}

\section{Advancement \& XP}
\label{sec:advancement}

Advancement in Fate's Edge reflects meaningful growth in capability and standing. XP is awarded for table-facing accomplishments, hard choices, and the dramatic friction that defines your story. Boons can be converted sparingly to accelerate growth.

\subsection{Awarding XP}
Choose a session pacing dial and stick to it for a campaign arc.
\begin{description}[leftmargin=1.5em, style=nextline]
  \item[Gritty:] 4--6 XP per session (slow burn).
  \item[Standard:] 6--10 XP per session (default pace).
  \item[Heroic:] 10--14 XP per session (fast growth).
\end{description}

\subsubsection{Session Awards (Guidelines)}
\begin{itemize}
  \item \textbf{Table Attendance:} +2 XP
  \item \textbf{Major Objective Reached:} +2--4 XP
  \item \textbf{Discovery or Lore Unlocked:} +1--2 XP
  \item \textbf{Hard Choice Embraced:} +1--2 XP
  \item \textbf{Complication Spotlight:} +1--3 XP
  \item \textbf{Bond/Flag Driven Play:} +1--2 XP
  \item \textbf{GM Curveball Award:} +0--3 XP
\end{itemize}

\subsubsection{Milestones}
\begin{itemize}
  \item At the conclusion of a major story arc, award \textbf{+8--12 XP to all players}.
  \item Grant \textbf{+2 XP} to one player for a signature moment of the arc.
\end{itemize}

\subsubsection{Boon Conversion}
Once per session, during downtime, a character may convert \textbf{2 Boons $\rightarrow$ 1 XP} (max \textbf{2 XP} via conversion per session). All normal Boon limits apply (hold 5; trim to 2 at scene end).

\subsection{Spending XP}
\label{sec:spendingxp}
\begin{description}[leftmargin=1.5em, style=nextline]
  \item[Attributes:] Cost $=$ \emph{new rating} $\times 3$. Downtime $=$ \emph{new rating} in days.
  \item[Skills:] Cost $=$ \emph{new level} $\times 2$. Downtime $=$ \emph{new level} in days.
  \item[On-Screen Followers:] Cost $=$ \emph{Cap}$^2$. Downtime $=$ 1--3 days to recruit and brief.
  \item[Off-Screen Assets:] Minor (4 XP, 1 day), Standard (8 XP, 1 week), Major (12 XP, 1 month).
  \item[Talents/Minor Assets:] As listed (e.g., Caster's Gift 6 XP; Patron's Symbol 4 XP).
\end{description}

\subsection*{Upkeep}

\textbf{Frequency:} Upkeep is paid once per Downtime period per asset.

\begin{itemize}
  \item \textbf{Option 1 — Efficient (Higher XP, Less Time)}
  \emph{Cost:} Pay Upkeep XP = $\max(1, \frac{\text{XP Acquisition Cost}}{3})$
  \emph{Time:} Minimal effort (e.g., sending a retainer, a quick inspection).

  \item \textbf{Option 2 — Intensive (Lower XP, More Time)}
  \emph{Cost:} Pay 1 XP
  \emph{Time:} A dedicated Downtime action of significant personal involvement (e.g., training a follower, repairing an asset).
\end{itemize}

\textbf{Failure:} If upkeep is not paid, the resource degrades:
\begin{itemize}
  \item \emph{Follower:} Becomes \textbf{Wary} (or \textbf{Seized} if already Wary).
  \item \emph{Asset:} Becomes \textbf{Neglected} (or \textbf{Compromised} if already Neglected).
\end{itemize}

\subsection{Rush Rule}
You may \emph{skip downtime} when buying or upgrading, but the GM creates a \textbf{Haste clock [4]}. If the clock fills, the new ability or asset carries flaws or narrative complications.

\subsection{Tiers of Reputation}
Reputation tiers reflect how the world responds to you.
\begin{description}[leftmargin=1.5em, style=nextline]
  \item[Tier I --- Rookie (0--40 XP):] Local reputation; prestige locked.
  \item[Tier II --- Seasoned (41--90 XP):] Regional notice; prestige abilities may unlock.
  \item[Tier III --- Veteran (91--150 XP):] National influence; second follower slot suggested.
  \item[Tier IV --- Paragon (151--220 XP):] Movers and shakers; rivals emerge to challenge.
  \item[Tier V --- Mythic (221+ XP):] Legendary status; kingdoms and cults respond.
\end{description}

\subsection{Advancement Notes}
\begin{itemize}
  \item Attribute cost scaling encourages diversification over single-stat spikes.
  \item Skill mastery yields tangible, fiction-first benefits.
  \item Prestige abilities (6+ XP) should be gated by narrative milestones or patron bargains.
\end{itemize}
