% Fate's Edge SRD --- Followers and Assets (Section 14)

\section{Followers and Assets}

Followers and Assets represent the lasting impact of your character's actions on the world. They are not abstractions or passive bonuses — they are people, resources, and structures that exist in the fiction and can act independently of your character.

Together, Followers and Assets allow characters to project influence beyond a single roll or scene, shaping the world through relationships, preparation, and legacy.

\subsection{Followers}

Followers are named NPCs who have a personal relationship with your character. They may be students, allies, retainers, agents, family members, or dependents. Followers act in the fiction and can grow, change, or be lost as the story progresses.

\subsubsection{What Followers Do}
\begin{itemize}
  \item \textbf{Assist} you during scenes through Bonds and Assists.
  \item \textbf{Act offscreen} to advance goals, gather information, or manage responsibilities.
  \item \textbf{Accumulate narrative weight} and may eventually become player characters.
\end{itemize}

Followers are not disposable tools. They have motives, limitations, and vulnerabilities, and the GM may place them at risk when the fiction demands it.

\subsubsection{Using Followers}
Followers typically influence play in three ways:
\begin{itemize}
  \item \textbf{Assists:} Granting +1 die or narrative positioning when they are present and able to help.
  \item \textbf{Bond Activation:} Spending a Boon to draw on trust, loyalty, or shared history for insight or support.
  \item \textbf{Offscreen Action:} Handling tasks that would otherwise demand the PC's time or attention.
\end{itemize}

\subsubsection{Follower States}
Instead of tracking numeric timers, note a follower's Loyalty and Fitness as one of three states each.

\textbf{Loyalty}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{State} & \textbf{Meaning} & \textbf{Mechanical Effect} \\
\midrule
Faithful & Committed, trustworthy & Normal assist, independent actions available. \\
Strained & Doubtful, pressured & Assist at –1 die; may refuse risky orders. \\
Broken & Betrayed or deserted & Follower leaves or becomes hostile. \\
\bottomrule
\end{tabular}

\textbf{Fitness}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{State} & \textbf{Meaning} & \textbf{Mechanical Effect} \\
\midrule
Ready & Uninjured, rested & Full function. \\
Hurt & Wounded, exhausted & Assist at –1 die; no independent actions. \\
Down & Incapacitated & Requires rescue or medical care. \\
\bottomrule
\end{tabular}

Change states when the fiction demands it (e.g., after a failed mission, a moral compromise, or taking damage). No rolls are required.

\subsubsection{Limits}
\begin{itemize}
  \item Followers do not replace player actions.
  \item Followers do not make rolls unless the GM explicitly calls for it.
  \item Followers cannot solve major conflicts without consequences.
\end{itemize}
When Followers succeed or fail, those outcomes become part of the story.

\subsection{Assets}

Assets are durable resources, positions, or structures under your character's control. They represent leverage — things that allow you to influence the world at scale.

Examples include:
\begin{itemize}
  \item A safehouse, laboratory, or fortress.
  \item A mercantile network or smuggling route.
  \item A political office or religious authority.
  \item A powerful relic, symbol, or institutional role.
\end{itemize}

\subsubsection{What Assets Do}
Assets:
\begin{itemize}
  \item Create advantages that persist beyond a single scene.
  \item Resolve problems that cannot be solved with a single roll.
  \item Shape the direction and scope of the campaign.
\end{itemize}
Assets are not passive bonuses. They matter only when invoked in the fiction.

\subsubsection{Asset Status}
Assets have a single Status that describes their overall health and usability.

\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Status} & \textbf{Meaning} & \textbf{Effect} \\
\midrule
Flourishing & Well-maintained, respected & +1 die on rolls involving the asset, +1 Yield. \\
Stable & Normal operation & No modifier. \\
Strained & Problems accumulating & –1 die on asset-related rolls, no Yield. \\
Collapsed & Non-functional & Requires quest or major downtime to restore. \\
\bottomrule
\end{tabular}

The Status worsens when the asset is neglected, attacked, or suffers a public failure. It improves through dedicated downtime actions or successful missions that benefit the asset.

\subsubsection{Activating Assets}
Most Assets require the expenditure of Boons to activate.

When you spend a Boon on an Asset, you may:
\begin{itemize}
  \item Advance or halt a Clock.
  \item Introduce or remove a complication.
  \item Secure resources, protection, or access.
  \item Influence events happening offscreen.
\end{itemize}
The GM determines the exact impact based on the scale of the Asset and the current situation.

\textbf{Activation Examples:}
\begin{itemize}
  \item \textit{``My spy network delays the patrol — tick the Alarm timer backward by 1.''}
  \item \textit{``The safehouse has a hidden exit — we escape at Controlled Position.''}
  \item \textit{``My trading post provides fresh horses — remove one Supply segment.''}
  \item \textit{``My guild contact suppresses the rumor — cancel an incoming complication.''}
\end{itemize}
The GM may veto an activation that exceeds the Asset's scope.

\subsection{Upkeep}

During downtime, each Asset and Follower you maintain requires one of the following:
\begin{itemize}
  \item \textbf{Attention (Intensive):} Spend a scene interacting with them (repairing, training, socializing).
  \item \textbf{Resources (Efficient):} Spend 1 XP (or a relevant String) to cover costs offscreen.
\end{itemize}
If you provide neither, the Asset or Follower degrades one step:
\begin{itemize}
  \item Assets: Flourishing $\rightarrow$ Stable $\rightarrow$ Strained $\rightarrow$ Collapsed.
  \item Followers: Faithful $\rightarrow$ Strained $\rightarrow$ Broken.
\end{itemize}
Two consecutive downtimes without upkeep collapses the Asset or loses the Follower entirely.

\subsection{Terrestrial Patrons}

Not all power comes from cosmic entities. Terrestrial Patrons are mortal factions or individuals — nobles, guildmasters, crime lords, generals, merchant princes — who offer tangible aid in exchange for obligation.

\subsubsection{How Terrestrial Patrons Work}
When you accept a favor from a Terrestrial Patron (access, resources, protection, an Asset), you mark +1 Obligation to that patron. Your Obligation Capacity (Spirit + Presence) applies normally, and exceeding it causes Fatigue exactly as with cosmic Patrons (see Section~\ref{sec:magic-system}).

Favors are typically Diamonds — leverage, rewards, or Assets that would otherwise cost XP. Examples:
\begin{itemize}
  \item A noble grants you a safehouse (Minor Asset) in exchange for a future service.
  \item A guildmaster provides legal immunity (String) for one heist.
  \item A crime lord lends you enforcers (Cap 2 followers) for a single mission.
\end{itemize}

\subsubsection{Clearing Obligation}
Unlike supernatural Patrons, terrestrial obligations are cleared by performing tasks for the patron. The GM sets the scope:
\begin{itemize}
  \item \textbf{Minor task} (deliver a message, pay a bribe): clears 1 Obligation.
  \item \textbf{Significant task} (escort a caravan, break a siege): clears 2 Obligation.
  \item \textbf{Major task} (assassinate a rival, broker a treaty): clears 3+ Obligation.
\end{itemize}
Each task becomes a story hook. Refusing a task when Obligation is full may turn the patron into a rival or enemy.

\subsubsection{Benefits of Terrestrial Patrons}
\begin{itemize}
  \item No XP cost for favors — only Obligation.
  \item Access to Assets, Strings, and Followers you could not otherwise afford.
  \item Patrons escalate with your tier: a Tier I patron offers local favors; a Tier III patron offers regional power.
\end{itemize}

\subsection{Player Trusts}

A Trust is a shared pool of XP that players contribute to, used to acquire and maintain party-owned Assets and Followers. Trusts represent a collective enterprise — a mercenary company, a trading guild, a spy network, a religious order.

\subsubsection{Forming a Trust}
Any player may contribute XP to the Trust pool. Contributions are voluntary and cannot be withdrawn. The Trust has no individual owner; spending from the pool requires majority agreement.

\subsubsection{What a Trust Can Buy}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Cost} & \textbf{Benefit} \\
\midrule
4 XP & Minor Asset (shared safehouse, workshop, contact network). \\
8 XP & Standard Asset (guild charter, spy ring, fortress lease). \\
9 XP & Follower (Cap 3, loyal to the Trust, not any single PC). \\
\bottomrule
\end{tabular}

Assets and Followers acquired through a Trust belong to the party as a whole. Any member may activate them (spending their own Boons).

\subsubsection{Upkeep and Delegation}
Each downtime, the Trust must pay upkeep for its Assets and Followers using the standard rules (1 XP or a scene per asset/follower). Alternatively, the Trust may delegate upkeep to a Cap 5 Follower:
\begin{itemize}
  \item The Cap 5 follower can manage up to 4 Assets (or any combination of Assets and Followers totaling 4 slots).
  \item The follower's own upkeep covers the delegated Assets — no separate XP cost.
  \item \textbf{Risk:} If the follower becomes Compromised (disloyal, captured, or killed), all delegated Assets become Neglected or are lost until the follower is restored.
\end{itemize}

\subsubsection{Dissolving a Trust}
If the party disbands, remaining XP in the pool is lost. Assets may be transferred to individual characters who pay the XP cost (or accept Obligation to a Terrestrial Patron to cover the difference).

\subsection{Followers vs. Assets vs. Patrons vs. Trusts}

\begin{tabular}{@{}p{2.5cm}p{2.5cm}p{2.5cm}p{2.5cm}p{2.5cm}@{}}
\toprule
 & \textbf{Followers} & \textbf{Assets} & \textbf{Terrestrial Patrons} & \textbf{Trusts} \\
\midrule
Nature & People & Resources / Structures & Mortal factions & Shared pool \\
Primary Role & Personal support & Strategic leverage & Favors for Obligation & Collective ownership \\
Activation & Assists, Bonds & Boon expenditure & Accepting favors & Majority vote \\
Risk & Emotional, narrative & Structural, political & Refusal turns hostile & Delegated follower compromise \\
Growth & Can become PCs & Expand or collapse & Escalate with tier & Assets/Followers persist \\
\bottomrule
\end{tabular}

\subsection{High-Tier Play}

As characters advance, Followers, Assets, Patrons, and Trusts become increasingly important. Higher-tier Talents and Assets consume Boons, forcing players to choose between immediate success and long-term influence.

At Tier III and above, your personal dice pool may be maxed, but a Cap 3 Follower adds +3 to assist rolls, and a Major Asset can shift the outcome of a scene without a roll.

Victory is no longer just surviving the roll. It is choosing which parts of the world move when you do.

