
% --- Fate's Edge SRD — Section 14: Range & Position Quick Reference ---
% Include this file from your main .tex with: \input{14-range-position-quickref.tex}

\section{Range \& Position Quick Reference}
\label{sec:range-position-qr}
\textit{This section condenses the guidance from \S8 into table form for fast table use.}

\subsection{Range Bands}
\begin{center}
\begin{tabular}{@{}llp{8.5cm}@{}}
\toprule
\textbf{Band} & \textbf{Default} & \textbf{Notes} \\
\midrule
Close & Touching & Grapples, knives; polearms from Near at Limited Effect. \\
Near  & Assumed  & Same room/street segment; quick step away. \\
Far   & Distant  & Same site but not in reach; needs time/route/long implement. \\
Absent& Off-screen & Outside scene frame; needs cut/travel to interact. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Movement (beats, not meters)}
\begin{itemize}
  \item \textbf{1 Move} shifts one band: Close$\leftrightarrow$Near or Near$\leftrightarrow$Far.
  \item \textbf{Dash (Action)} shifts two bands: Close$\rightarrow$Far or Far$\rightarrow$Close.
  \item \textbf{Terrain/Obstacles} may add +1 Move or require a test first.
  \item \textbf{Absent$\leftrightarrow$Scene} use a Travel clock [2--4] or a quick cut; under pressure treat as Risky.
\end{itemize}

\subsection{Melee Flag}
Mark two parties \textbf{in Melee} when they are in Near and directly engaged.
\begin{itemize}
  \item \textbf{Enter:} 1 Move from Near (Risky if under fire).
  \item \textbf{Leave:} 1 Move to break off; if threatened, test \emph{Disengage} at Risky. On Partial/Miss: soft consequence, remain in Melee.
\end{itemize}

\subsection{Targeting \& Tools (defaults)}
\begin{center}
\begin{tabular}{@{}lp{10.5cm}@{}}
\toprule
\textbf{Type} & \textbf{Default Reach / Off-band Rules} \\
\midrule
Melee/Touch & Close only (Near on long polearms at Limited Effect). \\
Thrown / Short magic & Near; at Far impose Limited Effect or +1 DV unless source says otherwise. \\
Firearms / Long bows / Line spells & Near and Far if stated; at off-band, reduce Effect or +1 DV. \\
Social & Near by default; at Far need voice/signal/proxy; at Absent need message/oath/setup. \\
No line of sight & Raise DV by +1, drop Effect one step, or require [REVEAL]/[MARK]. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Perception, Stealth, Cover}
\begin{itemize}
  \item \textbf{Notice:} Near normal; Far harder (DV +1 or Limited Effect) especially with noise/cover.
  \item \textbf{Hide:} Break LoS or move to Far; at Absent you're off-frame unless tracked.
  \item \textbf{Cover:} Improves Position or reduces consequence severity; quality/angle sets bump.
\end{itemize}

\subsection{Position (Controlled / Risky / Desperate)}
Position affects consequence severity on Partial/Miss (not DV).
\begin{itemize}
  \item \textbf{GM Spend (1 CP):} Shift Position one step worse for current action or apply a brief environmental shove.
  \item \textbf{Player Spend (1 Boon):} Shift Position one step better for current action or cancel a single 1-step GM shift.
  \item \textbf{Narrative Triggers (free):} Flanking, reinforcements, collapsing cover, superior leverage can move Position one step (say why).
\end{itemize}
\emph{Limits:} One Boon shift per action. Multiple forces don't stack beyond one step each way—resolve in order declared.

\subsection{GM Quick Cues}
\begin{itemize}
  \item Default to \textbf{Near} and \textbf{Risky/Standard}. Ask: “Do you need a beat to get there?” If yes, spend a Move.
  \item Call \textbf{Melee} when knives are actually in play; don't track hexes.
  \item Prefer one strong twist over nickel-and-diming with many petty penalties.
\end{itemize}
