% --- Fate's Edge SRD — Section 15: Travel Framework ---
\section{Travel Framework}
\label{sec:travel}

Journeys are more than distance; they are sequences of pressure, discovery, and consequence. The travel framework turns long roads into a series of dramatic legs without micromanaging miles.

\subsection{Legs and Travel Clocks}
\begin{itemize}
  \item Break a long journey into \textbf{legs}, each representing a significant change in terrain, danger, or political region.
  \item Each leg uses a \textbf{Travel Clock} (default size 4 segments).
  \item The clock advances through \textbf{progress actions}, \textbf{encounters}, or \textbf{scene resolution}.
  \item When the clock fills, the party arrives at the destination (or the next region). Unresolved complications carry forward.
\end{itemize}

\subsection{Roles on the Road}
Assign one role per character each leg. Roles may rotate each leg.
\begin{description}[leftmargin=1.5em, style=nextline]
  \item[Guide:] Sets course, reads signs, rolls navigation. On a success, tick the Travel Clock +1.
  \item[Scout:] Moves ahead; first contact with hazards or ambushes. On a success, the party cannot be surprised on the first encounter of the leg.
  \item[Quartermaster:] Manages supplies and encumbrance. On a success, prevent one Supply drop or recover 1 spent Supply.
  \item[Watch:] Maintains vigilance. On a success, the party gains +1 Position on the first defense roll of any ambush.
\end{description}
Each role action requires a roll (Attribute + appropriate skill, DV set by the GM). Failure may tick the Travel Clock backwards or introduce a complication.

\subsection{Encounters \& Events}
Every 1–2 segments filled, the GM introduces an encounter or hazard. At minimum, each leg should include \textbf{one meaningful encounter}.
\begin{itemize}
  \item \textbf{Discovery:} Ruins, strange omens, hidden sites.
  \item \textbf{Hazard:} Terrain challenge, illness, extreme weather (see \S\ref{sec:clocks} for hazard clocks).
  \item \textbf{Encounter:} Bandits, emissaries, beasts, or rival travelers.
  \item \textbf{Complication:} Bond tested, supplies strained, pursuit gains ground.
\end{itemize}
Encounters use the standard resolution loop (Position, DV, Outcome). On a failed travel roll, the GM may spend SB to introduce an encounter immediately.

\subsection{Supply \& Depletion}
Track shared party supplies with a \textbf{Supply Clock [4]}.
\begin{itemize}
  \item Tick the clock when the Quartermaster fails a roll, when the GM spends SB on scarcity, or when the party is forced to march without resupply.
  \item When the clock fills, the party runs out of a critical resource (food, water, fuel). Apply a Condition (\emph{Hungry}, \emph{Thirsty}, etc.) or mark 1 Fatigue per character.
  \item Resupply at settlements, successful foraging, or by spending a Cache token (see Asset rules).
\end{itemize}

\subsection{Optional: Regional Travel Decks}
If using the optional \textbf{Regional Decks} (see full rules), draw a card from the destination’s deck when the Travel Clock ticks. Suits suggest the nature of the encounter:
\begin{itemize}
  \item \textbf{Spades:} Place – a landmark, ruin, or terrain feature.
  \item \textbf{Hearts:} Actor – a traveler, patrol, or local guide.
  \item \textbf{Clubs:} Pressure – weather, injury, resource loss.
  \item \textbf{Diamonds:} Leverage – a shortcut, a cache, or a friendly face.
\end{itemize}
Highest rank may set the Travel Clock size (e.g., 4 for 2–5, 6 for 6–10, 8 for J–K, 10 for A).

\subsection{GM Quick Cues}
\begin{itemize}
  \item Travel should \textbf{advance story stakes}, not pause them. Use legs to foreshadow threats, deepen bonds, or reveal Patron omens.
  \item Let players showcase talents in their chosen roles.
  \item Mix discovery and hardship to keep legs tense but rewarding.
\end{itemize}