% --- Fate's Edge SRD — Section 16: Deck of Consequences (Optional) ---
\section{Deck of Consequences (Optional)}
\label{sec:deck-consequences}

The \textbf{Deck of Consequences} is a narrative tool for injecting drama, setbacks, and twists. It replaces or supplements GM fiat with randomized complications that remain thematically consistent. Its use is entirely optional.

\subsection{Deck Structure}
Use a standard 52‑card deck (jokers optional). Suits map to complication themes:

\begin{center}
\begin{longtable}{@{}llp{8cm}@{}}
\toprule
\textbf{Suit} & \textbf{Theme} & \textbf{Examples} \\
\midrule
Hearts & Social / Emotional & Betrayal, love triangle, family drama, ally under strain. \\
Clubs & Physical / Violent  & Ambush, wound, fatigue, weapon break. \\
Diamonds & Resources / Wealth & Supplies run low, theft, loss of funding, gear breaks. \\
Spades & Mystical / Supernatural & Omen, curse, patron demand, haunting. \\
\bottomrule
\end{longtable}
\end{center}

\subsection{Card Ranks \& Severity}
\begin{description}[leftmargin=1.5em, style=nextline]
  \item[Ace:] Scene‑altering twist; compels immediate response.
  \item[King/Queen/Jack:] Major complication with lasting effects.
  \item[10–8:] Moderate complication that reshapes the current scene.
  \item[7–5:] Minor complication; a nuisance that creates tension.
  \item[4–2:] Subtle complication or foreshadowing omen.
\end{description}

\subsection{Jokers (Optional)}
\begin{itemize}
  \item Red Joker: Catastrophic event (environmental collapse, patron intervention).
  \item Black Joker: Dark boon (immediate help, but with lasting cost or debt).
\end{itemize}

\subsection{When to Draw}
The GM may draw from the deck when:
\begin{itemize}
  \item A roll shows multiple 1s (e.g., three or more).
  \item The GM spends 2+ Story Beats at once.
  \item A major clock fills (scene or campaign clock).
  \item During travel (see \S\ref{sec:travel}) to generate an unexpected encounter.
\end{itemize}
Aim for 1–2 draws per session; more if the tone skews chaotic.

\subsection{Integration with Fiction}
\begin{itemize}
  \item Translate the raw card result into a fiction‑first complication. Do not apply mechanical penalties alone.
  \item The complication should build on what is already happening in the scene, not restart the story.
  \item Respect player agency – allow clever mitigation, but ensure the consequence lands.
\end{itemize}

\subsection{GM Quick Cues}
\begin{itemize}
  \item Use the deck to surprise yourself as well as the players.
  \item If a card doesn’t fit the current fiction, draw again or adjust the rank up/down by one.
  \item For campaign‑scale foreshadowing, draw a card at the end of a session and note its theme; introduce the complication in a later session.
\end{itemize}