% Fate's Edge SRD --- Travel Framework (Section 16)

\section{Travel Framework}

Journeys are more than distance; they are sequences of pressure, discovery, and consequence. The travel framework turns long roads into a series of dramatic legs without micromanaging miles.

\subsection{Legs and Travel Clocks}

\begin{itemize}
  \item Break a long journey into \textbf{legs}, each representing a significant change in terrain, danger, or political region.
  \item Each leg uses a \textbf{Travel Clock} (default size 4 segments).
  \item The timer advances through progress actions, encounters, or scene resolution.
  \item When the timer fills, the party arrives at the destination (or the next region). Unresolved complications carry forward.
\end{itemize}

\subsection{Roles on the Road}

Assign one role per character each leg. Roles may rotate each leg.

\begin{tabular}{@{}lp{3cm}p{7cm}@{}}
\toprule
\textbf{Role} & \textbf{Key Test} & \textbf{Benefit on Success} \\
\midrule
Guide & Wits + Survival & Tick the Travel Clock +1. \\
Scout & Wits + Stealth & Party cannot be surprised on the first encounter of the leg. \\
Quartermaster & Wits + Craft & Prevent one Supply drop or recover 1 spent Supply. \\
Watch & Presence + Notice & Party gains +1 Position on the first defense roll of any ambush. \\
\bottomrule
\end{tabular}

Each role action requires a roll (Attribute + appropriate skill, DV set by the GM). Failure may tick the Travel Clock backwards or introduce a complication.

\subsection{Encounters and Events}

Every 1–2 segments filled, the GM introduces an encounter or hazard. At minimum, each leg should include one meaningful encounter.

\begin{tabular}{@{}lp{4cm}p{6cm}@{}}
\toprule
\textbf{Type} & \textbf{Examples} & \textbf{Typical Resolution} \\
\midrule
Discovery & Ruins, strange omens, hidden sites & Investigation or exploration roll. \\
Hazard & Terrain challenge, illness, extreme weather & Survival or Endurance test; may use a hazard timer. \\
Encounter & Bandits, emissaries, beasts, rival travelers & Combat or social resolution. \\
Complication & Bond tested, supplies strained, pursuit gains ground & Roleplay or skill challenge. \\
\bottomrule
\end{tabular}

Encounters use the standard resolution loop (Position, DV, Outcome). On a failed travel roll, the GM may spend SB to introduce an encounter immediately.

\subsection{Supply and Depletion}

Track shared party supplies with a \textbf{Supply Clock [4]}.

\begin{itemize}
  \item Tick the timer when the Quartermaster fails a roll, when the GM spends SB on scarcity, or when the party is forced to march without resupply.
  \item When the timer fills, the party runs out of a critical resource (food, water, fuel). Apply a Condition (Hungry, Thirsty, etc.) or mark 1 Fatigue per character.
  \item Resupply at settlements, through successful foraging, or by spending a Cache token (see Asset rules).
\end{itemize}

\subsection{Optional: Regional Travel Decks}

If using the optional Regional Decks, draw a card from the destination's deck when the Travel Clock ticks. Suits suggest the nature of the encounter:

\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Suit} & \textbf{Theme} \\
\midrule
Spades & Place – a landmark, ruin, or terrain feature. \\
Hearts & Actor – a traveler, patrol, or local guide. \\
Clubs & Pressure – weather, injury, resource loss. \\
Diamonds & Leverage – a shortcut, a cache, or a friendly face. \\
\bottomrule
\end{tabular}

The highest rank may set the Travel Clock size (e.g., 4 for 2–5, 6 for 6–10, 8 for Jack–King, 10 for Ace).

\subsection{Weather and Terrain Modifiers}

The GM may apply modifiers based on conditions:

\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Condition} & \textbf{Effect} \\
\midrule
Difficult Terrain & Movement costs double; Charge/Skirmish require open ground (or Relentless talent). \\
Extreme Weather (rain, snow, heat) & First travel roll each leg is Desperate. \\
Hazardous Weather (storm, blizzard, sandstorm) & Travel Clock ticks +1 at leg start; navigation DV +1. \\
Favorable Weather & First travel roll each leg gains +1 Position. \\
\bottomrule
\end{tabular}

\subsection{Foraging and Subsistence}

A character may forage during a leg instead of taking a role action. Roll Wits + Survival (DV by terrain, typical 3–4).

\begin{itemize}
  \item \textbf{Clean Success:} Gain 1 Supply.
  \item \textbf{Partial:} Gain 1 Supply but mark 1 Fatigue (time or exposure).
  \item \textbf{Miss:} No Supply; GM gains 1 SB (scarce resources, dangerous plant, predator sign).
\end{itemize}

\subsection{GM Quick Cues}

\begin{itemize}
  \item Travel should advance story stakes, not pause them. Use legs to foreshadow threats, deepen bonds, or reveal Patron omens.
  \item Let players showcase talents in their chosen roles.
  \item Mix discovery and hardship to keep legs tense but rewarding.
  \item A failed travel roll should never result in ``nothing happens.'' Use SB to introduce an encounter or tick a hazard timer.
\end{itemize}
