% Fate's Edge SRD −−− Encounters (Section 17)

\section{Encounters}

All encounters — combat, social, environmental, magical — use the same core engine: Position, Effect, DV, Boons, and Story Beats (SB). Timers track progress toward resolution or the escalation of threats.

\subsection{Common Framework}

\begin{enumerate}
  ℹtem \textbf{State Intent & Approach.} What do you want to achieve, and how?
  ℹtem \textbf{Set Position & DV.} GM determines risk (Dominant/Controlled/Desperate) and difficulty (2–7+).
  ℹtem \textbf{Roll Attribute + Skill.} Each 6+ = success; each 1 = SB for the GM.
  ℹtem \textbf{Apply the Outcome Matrix.} Clean Success, Success & Cost, Partial, Miss.
  ℹtem \textbf{Spend SB.} GM introduces complications, shifts Position, or advances timers.
\end{enumerate}

\subsection{Timers in Encounters}

Timers track cumulative progress or increasing pressure.

\begin{itemize}
  ℹtem \textbf{Progress Timer } (4–8 segments): Fill to succeed (e.g., persuasion, escaping a maze).
  ℹtem \textbf{Threat Timer } (4–8 segments): Fill to trigger a disaster (e.g., reinforcements, structural collapse).
  ℹtem \textbf{Contest Timers:} Two timers — one for each side; first to fill wins.
\end{itemize}

\textbf{Ticking Timers:}
\begin{itemize}
  ℹtem Partial: 1 tick (GM discretion).
  ℹtem Clean Success: 2 ticks.
  ℹtem Miss: may tick opposition timers.
  ℹtem GM may spend SB to tick a timer.
\end{itemize}

\subsection{Social Encounters}

Use Persuasion Timers for extended negotiations, debates, or intimidation.

⫕section{Setting Up}
\begin{itemize}
  ℹtem \textbf{Target's Resolve:} Easy (4−segment), Moderate (6−segment), Hard (8−segment).
  ℹtem \textbf{Initial Position:} Based on relationship, leverage, and setting (Dominant = mutual trust, Desperate = hostile audience).
  ℹtem \textbf{DV:} Standard ladder (3–7+) with modifiers for cultural gaps, time pressure, etc.
\end{itemize}

⫕section{Non−Face Characters Can Help}
\begin{itemize}
  ℹtem \textbf{Setup:} Create an advantage — produce evidence, cause a distraction, or perform a display of force. Roll an appropriate skill. On success, grant the primary speaker +1 Position or +1 die.
  ℹtem \textbf{Assist:} Spend 1 Boon to give +1 die (narrated as a timely interjection or meaningful glance).
  ℹtem \textbf{Interrupt:} On a Miss, a non−face character can take the fallout (e.g., absorb a social rebuke) by marking 1 Fatigue.
\end{itemize}

⫕section{Persuasion Rolls}
Each distinct argument or lever advances the timer:
\begin{itemize}
  ℹtem \textbf{Clean Success:} +2 segments.
  ℹtem \textbf{Partial:} +1 segment; GM may spend SB for a complication (e.g., ``They agree, but you will owe a favor'').
  ℹtem \textbf{Miss:} No progress; GM may tick an Opposition Timer (4–6 segments).
\end{itemize}
When the Opposition Timer fills, the target hardens — Position worsens or the ask becomes impossible. When the Persuasion Timer fills, the target agrees as framed.

\subsection{Environmental Encounters}

⫕section{Terrain and Positioning}
\begin{itemize}
  ℹtem \textbf{Height Advantage:} +1 die or +1 Effect for attacks from above.
  ℹtem \textbf{Cover:} Partial cover (+1 DV to target you); Full cover (cannot be targeted directly). You may Hide behind full cover (Stealth vs Perception).
  ℹtem \textbf{Difficult Terrain:} Costs 2 Move actions to traverse one band (or a full action to Dash). Charge/Skirmish require open ground unless the Relentless talent is used.
\end{itemize}

⫕section{Hazards and Timers}
Active hazards (fire, flooding, toxic gas) are represented by Threat Timers:
\begin{itemize}
  ℹtem Tick the timer on missed rolls, GM SB spends, or at scene intervals.
  ℹtem When the timer fills, the hazard triggers (e.g., collapse, wave of fire).
  ℹtem Characters may attempt to slow or disable the hazard (e.g., Tinker to seal a gas leak, Magic to freeze spreading flames). Treat as a Progress Timer .
\end{itemize}

\textbf{Example: Collapsing Bridge [6]} — ticks on each Partial or Miss. At 3 segments, footing becomes Desperate; at 6, the bridge falls.

\subsection{Magical Encounters}

⫕section{Adjudicating Spells in Combat}
\begin{itemize}
  ℹtem \textbf{Free Caster (Spellcraft):} Weave then Cast (2 actions). Push (mark Fatigue) to compress to 1 action. Casting while Engaged worsens Position by one step unless the spell has [TOUCH] or [WARD].
  ℹtem \textbf{Runekeeper (Rites):} Invoke a Rite as 1 action; mark Obligation. Push It (+1 Obligation) for increased effect.
  ℹtem \textbf{Invoker (Symbols):} Ritual takes DV+1 rounds. Crack the Seal (instant) marks +2 Obligation (+3 for High Rites); Symbol becomes Compromised.
  ℹtem \textbf{Cantor (Songs):} Sing a Song (1 action, resolves at start of next turn). Push to resolve immediately (mark Fatigue + advance Corruption).
\end{itemize}

⫕section{Traps and Wards}
\begin{itemize}
  ℹtem \textbf{Discover:} Notice vs DV 4–6.
  ℹtem \textbf{Disable:} Contested roll vs trap's tick. Clean Success = neutralized; Partial = disable but generate SB; Miss = trap activates.
  ℹtem \textbf{Activation:} GM spends SB to trigger trap (damage, condition, timer advance).
  ℹtem \textbf{Wards:} Magical barriers with a DV (3–6). Crossing without dispelling triggers the effect and ticks any relevant Leash by the ward's DV.
\end{itemize}

⫕section{Summoning and Spirits}
\begin{itemize}
  ℹtem \textbf{Leash Capacity:} Cap + Spirit segments.
  ℹtem \textbf{Ticks:} Spirit takes Harm (1 per Harm level); acts against its nature (1); crosses a ward without dispelling (ticks = ward's DV); split focus (1).
  ℹtem \textbf{Leash fills:} Spirit acts once to its nature (potentially hostile) then departs.
  ℹtem \textbf{Boon Finesse:} Summoner may spend 1 Boon to clear 1 tick (once per round).
\end{itemize}

\subsection{Extended Rituals and Downtime Encounters}

Use a Progress Timer (4–12 segments). Each segment represents a scene or a block of time. Players describe their approach and roll an appropriate skill; success advances the timer. Complications (GM SB) may add segments, require rare components, or draw unwanted attention.

\subsection{Quick Reference Table}

\begin{tabular}{@{}lccc@{}}
⊤rule
\textbf{Encounter Type} & \textbf{DV (Default)} & \textbf{Typical Skills} & \textbf{Common Timers} \\
∣rule
Combat & 3 (Controlled) & Melee, Ranged, Athletics, Command & Mob Overwhelm [6], Hazard [4–8] \\
Social & 3 (Controlled) & Sway, Subterfuge, Insight, Command & Persuasion [4–8], Opposition [4] \\
Environmental & 3–4 (varies) & Athletics, Survival, Tinker, Stealth & Collapse [6], Flood [6], Chase [4] \\
Magical & 4–6 (specific) & Arcana, Lore, Spirit+Resolve & Leash [Cap+Spirit], Ritual [6–12] \\
⊥tomrule
\end{tabular}

\subsection{GM Reminders}

\begin{itemize}
  ℹtem Let non−specialists contribute — use Setup and Assist actions, and allow creative skill use.
  ℹtem Spend SB to escalate, not punish. A well−spent SB changes the situation, creating new choices.
  ℹtem Timers make progress visible. When a timer ticks, describe what changes in the fiction.
  ℹtem Follow the procedure for Armor & Fatigue: convert, apply remaining Harm (clears existing Fatigue), then add Fatigue. Minimum 1 Fatigue per hit.
\end{itemize}
