
% --- Fate's Edge SRD — Section 19: Tags & Outcomes Cheat Sheet (safe tables) ---
% Include this file from your main .tex with: \input{19-tags-cheatsheet.tex}

\section{Tags \& Outcomes Cheat Sheet}
\label{sec:tags-cheatsheet}

This section condenses common tags, outcomes, and modifiers into a one-glance reference. It supplements the full explanations in \S10.

\setlength{\extrarowheight}{2pt}

\subsection{Core Action Outcomes}
\begin{center}
\begin{longtable}{|p{2.4cm}|p{3.0cm}|p{8.0cm}|}
\hline
\textbf{Result} & \textbf{Meaning} & \textbf{GM/Player Guidance} \\ \hline
Hit (6) & Full success & Effect as stated, no new cost. \\ \hline
Partial (4--5) & Success + cost & Complication, worse Position, reduced Effect. \\ \hline
Miss (1--3) & Failure + cost & GM spends SB, introduces backlash/complication. \\ \hline
Critical (2x6) & Strong hit & Upgrade Effect or create new advantage. \\ \hline
\end{longtable}
\end{center}

\subsection{Common Tags}
\begin{center}
\begin{longtable}{|p{3.5cm}|p{10.0cm}|}
\hline
\textbf{Tag} & \textbf{Quick Reference Effect} \\ \hline
\texttt{[AREA]} & Hits all in zone; collateral risk. \\ \hline
\texttt{[ARMOR]} & Grants +1 Resistance against matching Harm. \\ \hline
\texttt{[BANISH]} & Expel Outsider; vs.\ Cap. On hit, it departs or weakens. \\ \hline
\texttt{[BOUND]} & Immobilizes target until broken. \\ \hline
\texttt{[COUNTER]} & Interrupt cast/ability in progress (DV by fiction). \\ \hline
\texttt{[DISPEL]} & End/suppress ongoing effect. \\ \hline
\texttt{[FOLLOW-UP]} & Trigger additional linked action; usually free if fiction permits. \\ \hline
\texttt{[HEAL]} & Remove or downgrade Harm/Condition. \\ \hline
\texttt{[ILLUSION]} & Fictional effect; resisted by disbelief or disruption. \\ \hline
\texttt{[MARK]} & Target acquires tracer; next action gains +1 die. \\ \hline
\texttt{[REVEAL]} & Expose hidden or cloaked subject. \\ \hline
\texttt{[TETHER]} & Bind two entities; strain if separated. \\ \hline
\texttt{[WARD]} & Block specific entity/domain; test Cap to cross. \\ \hline
\end{longtable}
\end{center}

\subsection{Position \& Effect Shorthands}
\begin{itemize}
  \item \textbf{Position:} Controlled (safe), Controlled (default), Desperate (big payoffs, big costs).
  \item \textbf{Effect:} Limited (weak), Standard (expected), Great (strong, extra reach).
  \item GM can spend 1 SB to worsen Position one step; Player can spend 1 Boon to improve Position one step.
\end{itemize}

\subsection{Backlash Reminders}
\begin{itemize}
  \item On Partial/Miss with Weave \& Cast, expect backlash colored by Element (fire burns, fate twists, etc.).
  \item Backlash does not add SB directly—it is a GM SB spend colored by the roll’s Element.
\end{itemize}

\subsection{Obligation Reminders}
\begin{itemize}
  \item Runekeeper Rite: +1 Obligation on completion, +1 more if Pushed.
  \item Invoker Ritual: Always +1 Obligation, even out of scene. Crack the Seal: +2/+3.
  \item Multi-symbol overload: carrying 4+ Symbols = +1 Obligation on first ritual each scene.
\end{itemize}

\subsection{Quick Costs}
\begin{itemize}
  \item \textbf{Caster’s Gift:} 2 XP. Required for Weave \& Cast.
  \item \textbf{Familiar/Thiasos:} 2 XP. Required for Patron’s Gift.
  \item \textbf{Codex:} 4 XP. Full Rite access; Obligations apply.
  \item \textbf{Patron’s Symbol:} 4 XP Minor Asset. Invoker ritual access.
\end{itemize}

\subsection{GM Quick Cues}
\begin{itemize}
  \item Default to Near + Controlled + Standard.
  \item Use tags as handles for rulings—if a move fits the tag, apply its shorthand.
  \item Complications escalate campaigns; use Clocks and the Deck of Consequences for follow-through.
\end{itemize}
