% Fate's Edge SRD −−− Crown Spread (Section 20)
% Optional campaign−framing tool

\section{Crown Spread (Optional)}

The Crown Spread is a campaign−framing tool that uses a spread of cards to establish the long arc of a story. It provides seeds for GMs and players to weave motifs, omens, and foreshadowed events. Its use is entirely optional.

\subsection{Setup}

\begin{itemize}
  ℹtem In Session Zero (or before the campaign begins), lay out 5–7 cards in a semicircle (the ``Crown''). Use either the Deck of Consequences or a standard card deck.
  ℹtem Each card anchors a motif, omen, or looming event.
  ℹtem Record the spread openly on a Campaign Sheet or digital log.
\end{itemize}

\subsection{Interpreting the Spread}

\begin{tabular}{@{}lp{10cm}@{}}
⊤rule
\textbf{Position} & \textbf{Meaning} \\
∣rule
Position 1 (Root) & The underlying tension or theme of the campaign. \\
Position 2 (Crest) & A key faction or patron influence that will rise. \\
Position 3 (Crown) & The climax image or major confrontation the arc builds toward. \\
Position 4 (Left Hand) & A bond, ally, or relationship that anchors play. \\
Position 5 (Right Hand) & A rival, betrayer, or challenger who pressures the party. \\
Positions 6 & 7 (Optional) & Expansions for setting−wide twists (environmental, mystical, or political). \\
⊥tomrule
\end{tabular}

\subsection{Using the Spread in Play}

\begin{itemize}
  ℹtem Each drawn card becomes a \textbf{Foreshadow Timer [4]} attached to its motif. Advance the timer when events lean toward that omen (e.g., when the GM spends SB on related complications, when a major scene involves that faction, or when a travel leg passes through an area tied to the motif).
  ℹtem When a Foreshadow Timer fills, the motif manifests concretely in play — a faction rises, a betrayer reveals themselves, a curse activates.
  ℹtem Tie Spread cards to Campaign Timers for pacing.
\end{itemize}

\subsection{Campaign Timer }

The Campaign Timer tracks rising stakes across the entire arc.

\begin{itemize}
  ℹtem Default size: \textbf{[8]} segments.
  ℹtem Advance the Campaign Timer when: multiple SB overflows in a session (e.g., three or more 1s on a single roll), when travel legs resolve with major cost (e.g., a character reaches Harm 2), or when Spread omens manifest concretely.
  ℹtem When the Campaign Timer fills, the Crown confrontation arrives. Play through its fallout as the campaign climax.
\end{itemize}

\subsection{Ending & Legacy}

\begin{itemize}
  ℹtem After the Crown confrontation resolves, hold an epilogue session.
  ℹtem Resolve any remaining Foreshadow Timers as epilogue vignettes.
  ℹtem Players may mark \textbf{Legacy Bonds} — new anchors for future campaigns or descendants.
\end{itemize}

\subsection{GM Quick Cues}

\begin{itemize}
  ℹtem The Spread is not a railroad — it foreshadows, not dictates.
  ℹtem Reinterpret cards liberally as play evolves; symbols matter more than literal events.
  ℹtem Remind players of their omens between arcs to build tension and payoff.
\end{itemize}

