% Fate's Edge SRD --- GM Guidelines (Section 22)

\section{GM Guidelines}

This chapter consolidates advice for running Fate's Edge smoothly, fairly, and with narrative focus. It covers session preparation, in-play adjudication, common pitfalls, and tools for managing cognitive load.

\subsection{Session Zero}

Before the campaign begins, establish the following with your players.

\subsubsection{Safety Tools}
\begin{itemize}
  \item \textbf{Lines:} Content that will not appear (e.g., no sexual violence, no harm to children).
  \item \textbf{Veils:} Content that happens off screen or fades to black.
  \item \textbf{X-Card:} Any player may stop a scene at any time without explanation.
\end{itemize}
Record Lines and Veils where everyone can see them. Keep an X-Card visible on the table.

\subsubsection{Tone and Expectations}
\begin{itemize}
  \item What is the campaign's primary genre? (intrigue, exploration, mercenary, horror, etc.)
  \item How lethal and how grim should the tone be?
  \item How much player vs. GM authority over the story?
  \item Expected session length and attendance.
\end{itemize}

\subsubsection{Campaign Frame (Optional)}
Consider using the Crown Spread (Section~\ref{sec:crown-spread}) to generate a shared campaign frame. Draw 5–7 cards to establish motifs, factions, omens, and a looming climax.

\subsection{The 30-Second Adjudication Loop}

When a player declares an action, resolve it with this loop:

\begin{enumerate}
  \item \textbf{Clarify intent and approach.} ``What do you want, and how?''
  \item \textbf{Set stakes and Position.} ``If it works, what changes? If it fails, what bites?'' Start Dominant/Standard unless fiction says otherwise.
  \item \textbf{Roll and read.} Count successes (6–9 = 1, 10 = 2); count 1s as SB.
  \item \textbf{Spend one beat well.} Cash SB on one memorable twist or tick a relevant Timer .
  \item \textbf{Push forward.} Describe how the fiction is now different. Ask, ``Who moves next?''
\end{enumerate}

\subsection{Defaults That Keep Things Moving}

When in doubt, use these defaults:

\begin{itemize}
  \item \textbf{Range:} Near (unless otherwise stated).
  \item \textbf{Position:} Controlled.
  \item \textbf{Effect:} Standard.
  \item \textbf{DV:} 3 (2 for trivial, 4 for hard, 5+ for extreme).
  \item \textbf{SB budget per scene:} 4 + highest character Tier. Unspent SB may carry over (max 4).
\end{itemize}

\subsection{When to Roll (and When Not To)}

\begin{itemize}
  \item \textbf{Roll when:} uncertainty exists \textbf{and} meaningful stakes are present \textbf{now}.
  \item \textbf{Do not roll when:}
  \begin{itemize}
    \item The outcome is certain (say ``Yes'').
    \item The action has no real cost or consequence.
    \item The player is fishing for Boons with repetitive, low-stakes rolls.
  \end{itemize}
  \item \textbf{Offer a choice or cost instead of a roll when:} the action is plausible but the player wants to avoid risk. ``You can do it, but it will take time (tick a timer)'' or ``You can do it, but you will mark 1 Fatigue.''
\end{itemize}

\subsection{Timers: The Three-Timer Rule}

At any given time, maintain at most \textbf{three active timers} in a scene:
\begin{enumerate}
  \item \textbf{Situation Timer :} immediate goal (persuasion, escape, research).
  \item \textbf{Counter-Clock:} opposition pressure (guards arriving, ritual completing).
  \item \textbf{Campaign Timer :} long-term front (faction rise, curse spreading).
\end{enumerate}
Merge or retire additional timers. Timers are visible; name them and tick them in front of players.

\subsection{Spending Story Beats (SB)}

\begin{itemize}
  \item \textbf{One strong beat is better than three tiny penalties.} Change the fiction, don't just add numbers.
  \item \textbf{Tie SB spends to visible outcomes:} a new foe appears, a path closes, a timer advances.
  \item \textbf{Surprise yourself.} If you don't know how a complication will resolve, spend SB and find out together.
\end{itemize}

\subsection{Balancing Encounters by Tier}

Fate's Edge does not use Challenge Ratings. Instead, consider the party's Tier and the enemy's Cap.

\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Party Tier} & \textbf{Typical Threat Cap} & \textbf{Notes} \\
\midrule
Tier I (0–40 XP) & Cap 1–2 & Single enemies; avoid area effects. \\
Tier II (41–90 XP) & Cap 2–3 & Lieutenants with minions. \\
Tier III (91–150 XP) & Cap 3–4 & Bosses with phase timers. \\
Tier IV (151–220 XP) & Cap 4–5 & Mythic threats; require preparation and assets. \\
\bottomrule
\end{tabular}

\subsubsection{Building an Enemy}
\begin{itemize}
  \item \textbf{Cap:} Determines dice pools, Harm output, and resistance.
  \item \textbf{Harm:} Typical Harm output = Cap (e.g., Cap 2 deals Harm 2 on a hit).
  \item \textbf{Defense DV:} 3 for standard foes, 4 for elites, 5+ for bosses.
  \item \textbf{Timers:} Bosses should have phase timers (6–10 segments) that change their behavior when filled.
\end{itemize}

\subsection{The Four Questions (When Stuck)}

Ask out loud:

\begin{enumerate}
  \item If this goes right, what changes? (Intent)
  \item If this goes wrong, what bites back? (Stakes)
  \item What single twist will make this memorable? (SB spend)
  \item Who moves next? (Momentum)
\end{enumerate}

\subsection{Common Pitfalls and Fixes}

\begin{tabular}{@{}lp{8cm}@{}}
\toprule
\textbf{Pitfall} & \textbf{Fix} \\
\midrule
Over-cranking SB (scenes feel punitive) & Halve your SB spends for a while; cash into visible timers instead of immediate penalties. \\
Timer sprawl & Merge redundant timers. Each active scene rarely needs more than 2–3. \\
Tag paralysis (player waits for perfect tag) & Paraphrase: ``Sounds like [VEIL]. DV 3. Want to roll?'' \\
Rules drift & If table memory conflicts with text, pick the ruling that keeps flow, then note to reconcile after play. \\
Player analysis paralysis & Offer two clear options and one risky third. ``You can A, B, or try something bold.'' \\
\bottomrule
\end{tabular}

\subsection{Handling Player Creativity}

\begin{itemize}
  \item \textbf{Say ``Yes, but'' not ``No.''} Offer a cost, a timer, or a complication.
  \item \textbf{Use Setup and Assist actions.} Let non-specialists contribute meaningfully.
  \item \textbf{Embrace unexpected solutions.} If a player bypasses a planned encounter, let it work — then escalate somewhere else.
  \item \textbf{Fail forward.} A missed roll should never stop the story; it should change direction.
\end{itemize}

\subsection{Session Checklist}

\subsubsection{Before Play}
\begin{itemize}
  \item Review previous session's fallout and unresolved timers.
  \item Update faction and front timers.
  \item Prepare 2–3 flexible scenes (not scripts).
  \item Set the initial Position and DV for the opening scene.
  \item Remind players of safety tools.
\end{itemize}

\subsubsection{During Play}
\begin{itemize}
  \item Use the 30-second loop.
  \item Spend one strong beat per roll (not many small penalties).
  \item Keep timers visible and tick them openly.
  \item Rotate spotlight among players.
  \item When you don't know a rule, make a ruling and note it for later.
\end{itemize}

\subsubsection{After Play}
\begin{itemize}
  \item Award XP (Section~\ref{sec:advancement}).
  \item Advance or clear active timers.
  \item Note rulings to reconcile with rules text.
  \item Ask players: one thing they loved, one thing they want more of.
\end{itemize}

\subsection{Rookie GM Comfort Dials}

If you are new to Fate's Edge, start with these training wheels:

\begin{itemize}
  \item \textbf{Soft SB:} For your first 2 sessions, cap each roll's SB spend to 1–2 unless it is a set-piece.
  \item \textbf{Visible Timers:} Put all timers on the table. Name them aloud. Tick them in ink.
  \item \textbf{Tag Cards:} Print one-liners for frequently used tags ([WARD], [BANISH], [COUNTER]). Hand them out when a power is active.
  \item \textbf{One Move, One Sentence:} Every ruling should end with one sentence that states the new situation.
\end{itemize}

\subsection{Design Guardrails (for Consistency)}

These principles guide rulings and homebrew:

\begin{itemize}
  \item \textbf{Narrative Primacy:} Mechanics serve story, not replace it.
  \item \textbf{Risk as Drama:} Every roll carries potential for triumph and complication.
  \item \textbf{Meaningful Growth:} XP changes characters and the world.
  \item \textbf{Consequence Weight:} Choices ripple outward; nothing is free.
  \item \textbf{Fail Forward:} Misses fuel Boons; 1s become SB.
\end{itemize}

\subsection{Final Advice}

The rules are rails you lay just ahead of the train. Do not let perfect knowledge of the system slow the table. When in doubt, follow the fiction, set a Position and DV, and roll. The game is designed to carry you.

