% Fate's Edge SRD --- Appendix: Example Characters

\section{Example Characters}

These Tier I (32 XP) characters illustrate core builds for different playstyles. Each includes attributes, skills, talents, equipment, a bond, and a complication.

\subsection{Kaelen — Aeler Initiate (Warrior/Runekeeper)}

\textit{A young vent-prior's apprentice who left the underways to prove that stone-law can walk the surface world without breaking.}

\subsubsection*{Attributes}
\begin{tabular}{@{}ll@{}}
Body 3 & Wits 2 \\
Spirit 3 & Presence 2 \\
\end{tabular}

\subsubsection*{Skills}
Melee 2, Athletics 1, Endurance 1, Lore 1, Craft 1, Survival 1

\subsubsection*{Talents (12 XP total)}
\begin{itemize}
  \item \textbf{Familiar (Thiasos) (2 XP)} — A small iron bell that hums when stone is near.
  \item \textbf{Codex (4 XP)} — Sealed with a keystone rune; contains the Rites of the Sealed Gate.
  \item \textbf{Weapon Master (5 XP)} — +2 dice with warhammer; once/scene +1 Effect.
  \item \textbf{Effortless Charge (4 XP)} — Ignore Fatigue cost of Charge once/scene.
\end{itemize}

\subsubsection*{Equipment}
\begin{itemize}
  \item Warhammer (Medium melee, Harm 2, +1d at Close, +2d at Near).
  \item Iron Half-Plate (Medium armor, 1 Harm $\rightarrow$ 1 Fatigue, 2 Harm $\rightarrow$ 1 Fatigue).
  \item Codex (iron-bound slate).
  \item Breath-counter bell.
\end{itemize}

\subsubsection*{Bond}
``I owe Vent-Prior Orla a debt of stone and silence. She taught me that the mountain listens even when we do not speak.''

\subsubsection*{Complication}
\textbf{Stone-Press Susceptible:} In deep underground spaces, any failed roll ticks a personal Hallucination Clock [4].

\subsubsection*{Rites Known}
\begin{itemize}
  \item \textbf{Sealed Threshold} (Low) — Ward a doorway; unauthorized crossing requires Resolve test.
  \item \textbf{Oath-Ward} (Low) — Chalk circle wards a small area vs intrusion.
\end{itemize}

\subsection{Lyra — Vilikari Rover (Rogue/Scout)}

\textit{A border-runner who learned to read two ledgers before she could write her name, now carries secrets instead of cargo.}

\subsubsection*{Attributes}
\begin{tabular}{@{}ll@{}}
Body 2 & Wits 4 \\
Spirit 2 & Presence 3 \\
\end{tabular}

\subsubsection*{Skills}
Stealth 2, Subterfuge 2, Sway 2, Investigation 1, Tinker 1, Insight 1

\subsubsection*{Talents (12 XP total)}
\begin{itemize}
  \item \textbf{Keen Senses (2 XP)} — +1 die on perception checks to spot danger or hidden details.
  \item \textbf{Trackless Step (2 XP)} — Once/scene, leave no trail for the remainder of the scene.
  \item \textbf{Silk Palms (2 XP)} — When you bribe a minor official, gain temporary String \textit{Bought Silence}.
  \item \textbf{Backstab (6 XP)} — When attacking from stealth, deal +1 Harm and ignore 1 point of Armor.
\end{itemize}

\subsubsection*{Equipment}
\begin{itemize}
  \item Paired daggers (Light melee, Harm 1, +2d at Close, +1d at Near).
  \item Padded leather (Light armor, 1 Harm $\rightarrow$ 1 Fatigue, 2 Harm $\rightarrow$ 1 Harm + 1 Fatigue).
  \item Lockpicks (Tinker kit).
  \item False manifest (forged papers).
\end{itemize}

\subsubsection*{Bond}
``Rurik the fence saved me from a customs trap. He expects me to return the favor when the ledger calls.''

\subsubsection*{Complication}
\textbf{Two-Ledger Debt:} You keep a second ledger of favors owed to Vilikari smugglers. The GM may once per session tick a \textit{Favor Due} timer [4] when you refuse a questionable job.

\subsection{Sera — Lethai-thora Exile (Caster)}

\textit{A former archivist who touched a forbidden scroll and now hears whispers from the space between written words.}

\subsubsection*{Attributes}
\begin{tabular}{@{}ll@{}}
Body 2 & Wits 4 \\
Spirit 3 & Presence 2 \\
\end{tabular}

\subsubsection*{Skills}
Arcana 2, Lore 3, Insight 2, Investigation 1, Deception 1

\subsubsection*{Talents (10 XP total)}
\begin{itemize}
  \item \textbf{Spellcraft (6 XP)} — Enables freeform casting (Weave + Cast).
  \item \textbf{Long Memory (3 XP)} — +1 die to Lore and Insight rolls involving historical knowledge.
  \item \textbf{Cold Reading (3 XP)} — +1 die to Sway and Insight when observing before acting.
\end{itemize}

\subsubsection*{Equipment}
\begin{itemize}
  \item Quarterstaff (Light melee, Harm 1, +2d at Close, can Defend as action).
  \item Scholar's robes (no armor).
  \item Sealed scroll case (contains one forbidden fragment).
  \item Brass inkpot (focus for [REVEAL] effects).
\end{itemize}

\subsubsection*{Bond}
``Keeper Lian hid me when the Censorate came. Her silence protects me; my voice will one day protect her secret.''

\subsubsection*{Complication}
\textbf{Whisper-Threshold:} Whenever you cast a spell that generates two or more SB, you hear a voice from the Ninth Scroll. The GM gains a String: \textit{The Archive Knows Your Name}.

\subsubsection*{Signature Spells}
\begin{itemize}
  \item \textbf{Whisper-Light} ([FIRE][ILLUMINATE]) — DV 3, one action. Small, silent light.
  \item \textbf{Memory's Veil} ([WATER][VEIL]) — DV 3, two actions. Obscure one small object or door; –1 die to Notice.
  \item \textbf{Scroll's Rebuke} ([FATE][STRIKE]) — DV 3, two actions. Ranged Harm 1 (psychic), ignores mundane armor.
\end{itemize}


