\documentclass[11pt]{article}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{tcolorbox}

\geometry{margin=1in}
\setlength{\parskip}{6pt}

\begin{document}

\title{Fate's Edge: Player's Guide Primer}
\author{}
\date{}
\maketitle

\tableofcontents

\section{Welcome to Fate's Edge}
Fate's Edge is a world where choices ripple outward, magic is perilous, and cultures define more than appearance—they shape destiny. 
This primer offers the foundation: lands, tongues, gods, and peoples. It also introduces the three XP paths that shape characters in play.

\section{The Northern Shore \& Amaranthine Sea}
The heart of adventure lies along the Amaranthine Sea and its northern shore, a mosaic of kingdoms, republics, and wilderness.

\section{Major Regions}
\subsection*{Post–Utar / Upper Amaranthine Basin — “Shards of an Empire”}
\textbf{Tagline:} River thrones, hill banners, and city-states stitched together by roads older than their rulers.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Successor realms ring the Amaranthine and Belworth systems: old Utar law persists in ledgers, not loyalties.
  \item Two great confessions—Everflame and Light—compete with local cults and dwarven compacts.
  \item Roads, ferries, and coin-houses matter as much as castles; mercenary companies thrive.
\end{itemize}

\begin{description}[leftmargin=2.2cm]
  \item[Ecktoria] Rump of the Utar heartland: gladiator arenas, Everflame processions, marble in need of mortar. A wealthy capital city rides habits of empire while petty principalities fill the hinterland. Exports ceremony and coin, imports grain and peace.

  \item[Acasia] Failed province of ruined prefectures and opportunist crowns; strongmen rule mile by mile. \emph{Silkstrand} is the bright anomaly—oligarch merchants, bravos, and blue-water trade. Caravans hug safe corridors; everyone else pays someone.

  \item[Vhasia] Old union of blood-houses and dukedoms; pride hardens into rivalry. Lence keeps the royal fiction alive; Vhaston, Eastwell, and Ballafield pull in different directions. Oath, marriage, and musters remain the coin of governance.

  \item[Viterra] Highlands, fens, and dales under a lawful temperament and long memory. River tolls are counted to the nail; the King’s/Queen’s Highway customs are sacred. Knights of the Dawn keep roads honest; counting houses keep everyone else so.

  \item[Ubral] Foothill clans and hill dwarves knit by feud law and fast marches. Hit-and-run war is a craft; hospitality is binding once given. Trade runs in cattle, iron blooms, and guides who know every sheep-track and scree path.

  \item[Thepyrgos] Astroegro’s last lantern: universities, orators, ateliers, and archives amid kraken-haunted lore. High-elf enclaves maintain salons and schools; harbors hum with debate as much as ships. Beauty, rhetoric, and contracts share a bench.

  \item[Mistlands] Dwarven “protectorate” of peat farms and fog, where waystones and ward-mounds hold back stranger things. Payden’s Port is the hinge; fae incursions are a seasonal line item. Human steads prosper—inside agreed boundaries.

  \item[Theona] Three tight islands of hedgerows, marble cuts, and yeoman plots; duchies argue quietly while fields work loudly. Claims from the mainland wax and wane; parish charters outlast dukes. Grain, stone, and stubbornness export well.

  \item[Kahfagia] Maritime oligarchy on the western quarter: army, navy, marines, clergy, civics, and privateers share the helm. Tempest rites bless pilots and warlocks alike; charter ports project power. Finished goods out, prize law in.

  \item[Galanina \& Fharan Lands] Cedar coasts and caravan capitals: Sidhi courts, Fharan trade cities, and shrine networks that make deserts crossable. Silk and spice routes braid cultures; litigies and ledgers travel with caravans.

  \item[Haayr Peninsula] Mountain-ridged hinge between seas; palimpsest of empires. Ancient Hyro stones, Ashaani ruins, and current Kuvani prefectures overlap. Whoever holds the passes writes the tariffs—until the next holder does.
\end{description}

\section{Travelers' Map (Optional Placeholder)}

\noindent
\textit{This section is reserved for a visual map of the Amaranthine Sea region and surrounding lands. Until the map is finalized and included with:}  
\begin{center}
  \texttt{\textbackslash includegraphics[width=\textbackslash textwidth]\{map\_fatesedge.pdf\}}
\end{center}

\subsection*{Orientation Notes}
Until the map is prepared, use the following compass-points summary to orient play:  

\begin{description}[leftmargin=2cm]
  \item[North] Mistlands and Payden’s Port; dwarf protectorates and Fae-haunted fogs.  
  \item[East] Valewood, Amedell, and the Alberriden Sea; homeland of Wood Elves and Halflings.  
  \item[South] Amaranthine Sea and the broken provinces of Utar; Ecktoria, Vhasia, and Viterra dominate.  
  \item[West] Kahfagia across the Titan’s Throat, naval empire of storms and krakens.  
\end{description}

\subsection*{Major Routes}
\begin{itemize}
  \item \textbf{The Belworth Corridor} — Main artery between Vhasia and Viterra, heavily fortified.  
  \item \textbf{The Violet Steppes} — Wide-open plains north of Utar, cradle of Ykrul incursions.  
  \item \textbf{The Way of Silk} — Southern caravan trade linking Fhara, Kuvani, and Dhahara to northern ports.  
  \item \textbf{The Titan’s Throat} — The narrow strait where the Amaranthine meets the Great Sea; lifeline for Kahfagian fleets.  
\end{itemize}

\subsection*{Use at the Table}
\begin{itemize}
  \item \textbf{As a GM:} Drop pins for encounters, migrations, or rumors to create a sense of travel.  
  \item \textbf{As a Player:} Use the cardinal notes above to keep track of where your character has journeyed.  
  \item \textbf{At Character Creation:} Mark one home region and one place of past travel to immediately ground your PC in the world.  
\end{itemize}


\section{Languages of the Lands}
Language is both bridge and barrier. Most folk speak \textbf{Common}, a dwarven-forged trade tongue. 
Yet each culture prizes its own tongue:
\begin{itemize}
  \item \textbf{Low Utaran} — the speech of peasants and soldiers.  
  \item \textbf{High Utaran} — the scholar’s and priest’s tongue.  
  \item \textbf{Vhasian Dialects} — Lence’s courtly form vs. Vhaston’s trader’s patter.  
  \item \textbf{Viterran Plain Speech} — blunt, pragmatic, precise.  
  \item \textbf{Dwarven (Aeler)} — guttural, clipped, rich in trade terms.  
  \item \textbf{Elven (Lethai)} — ancient, fluid, contextual.  
  \item \textbf{Tulkani Tongue} — lilting, mobile, woven with shadow-cant.  
  \item \textbf{Kuvani Speech} — sharp, consonant-heavy, linked to steppe songs.  
\end{itemize}

\section{Far-Off Lands (Brief Gazetteer)}

\\subsection*{Akilan — “Continent of Sun, Stone, and River”}
\textbf{Tagline:} Cradle of empires, where river silt feeds scholar-kings and savanna caravans carry the weight of gold.\\
\textbf{Snapshot:}
\begin{itemize}
  \item A vast continent of monsoon coasts, river-fed empires, and savanna caravan routes linking the inland riches to blue-water ports.
  \item Power centers range from the bureaucratic machine of Oshiira and the divine mandate of Taharka to the mercantile hinges of Sekogo and the stone-kept secrets of the Ngomebe.
  \item Exports: Gold, spices, hardwoods, ivory, trained scribes, and monsoon pilots. Imports: Wrought steel, northern grain, books, and mercenary contracts.
  \item Languages: Oshiiran, Taharkan, Sekogo trade-pidgin, Ngombe stone-cant, and ancient Ashaani liturgical scripts.
  \item Hooks: An Oshiiran prefect hires guards for a census in the restless Crimson Basin; a Taharkan convoy master needs discreet muscle to bypass a Sekogo quay syndicate's ``weighing fee''; a Ngombe mason-guild offers strong coin for the recovery of a stolen resonance-crystal from a sunken temple.
\end{itemize}

\begin{center}
\begin{tabular}{|p{3cm}|p{3.5cm}|p{3.5cm}|p{5.5cm}|}
\hline
\textbf{Nation} & \textbf{Government/Identity} & \textbf{Exports} & \textbf{Adventure Hook} \\
\hline
\textbf{Oshiira} & Scholar-bureaucracy; riverine prefectures ruled by archivists and census-takers. & Scribes, law codes, jade seals. & A missing tax-roll imperils a governor’s mandate. \\
\hline
\textbf{Taharka} & Theocratic dynasty claiming divine sun-mandate; convoy masters and priest-kings. & Gold dust, ritual bronze, caravan oxen. & A rival sun-priest raises a heretical banner; adventurers hired to suppress it. \\
\hline
\textbf{Sekogo} & Mercantile hinge; coastal syndicates and guild councils dominate trade. & Salt, kola, indigo, mercantile services. & A syndicate clerk offers coin if smugglers vanish before tomorrow’s ledger is read. \\
\hline
\textbf{Ngomebe} & Mountain stone-keepers; resonance-stone masons and secret archivists. & Resonance crystals, cut stone, elephant ivory. & A fallen vault floods a valley with strange harmonics; escorts needed to reseal it. \\
\hline
\end{tabular}
\end{center}

\subsection*{Sihai — “Thousand Terraces, Ten Thousand Seals”}
\textbf{Tagline:} River dynasties, scholar-magistrates, and calendars that bind empires.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Heartland river plains ringed by mountains; irrigation guilds and terrace clans feed sprawling capitals.
  \item Bureaucracy of examination-seals; mandarins arbitrate clan feuds, salt monopolies, and spirit-rights.
  \item Exports: tea, silk, paper, porcelain, ink-stones. Imports: horses, fine ore, foreign mercenaries.
  \item Languages: High Sihai (court/ritual), Low Sihai (market); Common on the frontier.
  \item Hooks: A counterfeit seal-ring undermines a prefecture; a dragon-flood threatens the grain fleet unless taboo stones are moved.
\end{itemize}

\subsection*{Dhahara — “Peninsula of Sky and Memory”}
\textbf{Tagline:} Monsoon coasts and dust kingdoms where the god of the open sky still judges men by their banners.\\
\textbf{Snapshot:}
\begin{itemize}
  \item A long peninsula of river belts, thorn-steppe, and desert basins; caravan roads link upland passes to warm-water ports.
  \item \textbf{North}: the \emph{Himdal Marches}—Aeler-administered marches where old Dhaharan clans and ancient Kuvani blood intermix under fortress law. 
  \item \textbf{South}: Kahfagian charter-ports and “concessions” dominate deep-harbor trade; local courts retain inland rule but answer to quay ledgers.
  \item The majority faith venerates \textbf{Dhahar}, Lord of the Sky and War; coastal cities layer Tempest rites for safe harbors; caravan shrines keep the Way of Silk/Spice.
  \item The \textbf{Temple of Sealing} (old Dhaharan trust) and its sundered orders: a public fraternal lineage that treats the relic as legend, and a hardline cloister that keeps the true warnings.
  \item Exports: war-mounts, fine cotton, dyes, pepper, frankincense/myrrh resins, carved jade, caravan services. Imports: northern steel, coin-credit, ship timbers, learned clerks.
  \item Languages: Dhaharan (court and caravan), Himdali Aeler (march law), Kahfagian port-cant; Common on the roads.
  \item Hooks: A sky-feast tourney turns bloody after a banner is defamed; a concession-port freezes a convoy over “storm-damage fees”; a Himdal prefect seeks neutral arbiters for a clan water-rights moot.
\end{itemize}

\begin{center}
\begin{tabular}{|p{3.4cm}|p{4.1cm}|p{4.2cm}|p{5.2cm}|}
\hline
\textbf{Polity / Region} & \textbf{Government / Identity} & \textbf{Known For} & \textbf{Adventure Hook} \\
\hline
\textbf{Himdal Marches} (Northern uplands) & Aeler praefects and march-lords over mixed Dhaharan--Kuvani clans; fortress law, road tolls, census writs. & Border horse, hill-forts, pass control, levy-archers. & A census stone goes missing; three clans blame each other—recover it before the prefect levies a punitive tithe. \\
\hline
\textbf{Upper Dhahara} (Interior basins) & Clan emirs and sultanates bound by sky-oaths; water-right councils arbitrate grazing and wells. & Pepper gardens, incense groves, banner-tourneys, caravan guards. & A sky-oath duel is fixed; prove it without shaming both houses and starting a banner-feud. \\
\hline
\textbf{Lower Dhahara} (Monsoon belt) & City councils under hereditary malikates; Dhahar rites blended with harbor offerings. & Cotton looms, dye-vats, river-quays, monsoon pilots. & Monsoon charts are forged; expose the quay-syndicate before a fleet founders on a lee shore. \\
\hline
\textbf{Kahfagian Concessions} (Deep ports) & Charter harbormasters answer to Kahfagia’s oligarchy; local courts kept at arm’s length. & Customs houses, bonded warehouses, privateers’ brokers. & A “storm clause” traps a whole convoy in bonded storage; slip the goods out legally—or cleverly. \\
\hline
\textbf{The Temple Districts} (Sealing sites) & Public \emph{Brotherhood of the Seal} (ritual, civic charity) and hidden \emph{Custodians} (true-keeping, warding). & Jade pedestals, mnemonic chants, sealed vaults, desert hermitages. & A novice dies carrying a half-burned warning; follow the route before the desert swallows the rest. \\
\hline
\textbf{Kuvani Foothills} (North/east skirts) & Kuvani-descended riders under Dhaharan law; truce-markets and feud seasons. & Steppe ponies, horn bows, caravan outriders. & Feud-season spills onto the Silk road; negotiate a truce so the season’s last convoy can pass. \\
\hline
\end{tabular}
\end{center}

\paragraph{Faith and Festivals.}
Sky-feasts mark the first clear night after the rains; banners are judged for truth and craft. Breaking a sworn banner-oath is a public crime \emph{and} a sacrilege. Caravan gods are honored at waystones; at ports, Tempest priests bless departures while Dhahar’s kola-ash is thrown to the wind.

\paragraph{Law on the Road.}
Waystones guarantee right-of-passage to anointed caravans: harming them invokes both clan reprisals and march law. Quay charters cannot void waystone protections inland—merchants hire witnesses who know the difference.

\paragraph{Orders of the Seal.}
The public Brotherhood keeps libraries, schools, and civic rites around old foundations; the hidden Custodians maintain wards, train “rememberers,” and quietly move relics when politics lean dangerous. Both claim descent from the same vow; neither admits the other is wrong.

\paragraph{Trade Patterns.}
Caravans move pepper, resin, cotton, and carved jade north to river deltas and west to the Amrathian; iron, coin-credit, and timber flow back. In lean monsoons, resin and pilots fetch double; in fat ones, horses and guards do.

\subsection*{Nihon — “Isles of Quiet Steel”}
\textbf{Tagline:} Mountain isles, sea-lords, and sword schools that cut the wind.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Volcanic spines and cedar coasts; domains sworn to sea-lords and shrine courts.
  \item Duels of lineage styles; smiths fold storm and silence into steel.
  \item Exports: tempered blades, lacquer, pearl, dried fish. Imports: iron ore, horses, rare texts.
  \item Languages: Nihonese; shrine-cant among priestesses; Common in treaty ports.
  \item Hooks: A school’s secret kata is stolen before midsummer bouts; a reef-chart goes missing as typhoon bells ring.
\end{itemize}

\subsection*{Fharan Lands — “Sands of Frankincense and Fire”}
\textbf{Tagline:} Desert caravans, incense kingdoms, and star-watching citadels.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Oases string trade across red dunes; coastal incense cities fund desert fortresses.
  \item Tribal confederacies swear to city imams and salt princes; hospitality sacred, vendetta older still.
  \item Exports: frankincense, myrrh, glass, astrolabes. Imports: grain, timber, trained engineers.
  \item Languages: Fharan tongues; calculator cants in observatories; Common in caravanserais.
  \item Hooks: A sacred well is poisoned on the eve of treaty; a glassworks seeks protection from sand-cults.
\end{itemize}

\subsection*{Ashaan — “Thrones Under Sand”}
\textbf{Tagline:} River-lords, sun-temples, and tombs that still pay taxes.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Black silt kingdoms along a long river; green flood-bands bracketed by stone and desert.
  \item Viziers balance temple coffers and irrigation; desert cavalry patrol necropolis roads.
  \item Exports: grain, linen, natron, gold leaf, papyrus. Imports: timber, tin, mercenaries.
  \item Languages: Old Ashaani (liturgical), Bazaar Ashan (trade); Common with foreigners.
  \item Hooks: A tomb-audit goes lethal as ledgers vanish; a floodgate curse locks the river before planting.
\end{itemize}

\subsection*{Ayohkhan — “Emerald Archipelago of Smoke and Spice”}
\textbf{Tagline:} River-delta kingdoms, jungle thrones, and trade winds that choose their own kings.\\
\textbf{Snapshot:}
\begin{itemize}
  \item Mangrove deltas and mountain jungles; stilt cities rule estuaries while hill citadels tax passes.
  \item Spice unions sway thrones; monsoon fleets leapfrog isles with pilot-priests.
  \item Exports: cinnamon, clove, camphor, bird-nest delicacies. Imports: iron, coin, scribes.
  \item Languages: Ayohkhan speech families; pilot litanies at sea; Common in entrepôts.
  \item Hooks: A monsoon calendar is forged to rig spice futures; hill rebels seize the only pass before harvest.
\end{itemize}

\section{The Gods and Powers}
Religion is both faith and faction. Each power gives rise to cults, orders, and schisms:
\begin{itemize}
  \item \textbf{Everflame} — Adar, Odur, Akilesh as facets of one holy fire. Seat of the Church in Ecktoria.  
  \item \textbf{Light} — Evangelical offshoot of the Everflame, rooted in Viterra.  
  \item \textbf{Khemesh \& Ráeyn} — Lord of Krakens and Lady of Storms, twin patrons of Kahfagia.  
  \item \textbf{Sky-Spirit of the Ykrul} — unnamed, shamanic, tied to omens.  
  \item \textbf{Dhahar} — war-god of the southern peninsula that bears his name.  
  \item \textbf{Ikasha, She Who Sleeps Within the Shadow} — Tulkani matron, whispered more than preached.  
  \item \textbf{Old Ancestors} — Dwarven reverence of kin, blood, and the Stone.  
\end{itemize}

\section{Cultures at a Glance}

\section{Humans}
Adaptable, ambitious, and diverse. Found across every province and sea.  
Affinity: \emph{Versatility}.  

\subsection{Sub-Cultures}
\subsubsection{Viterrans}
Highland pragmatists and fen traders. Known for their \emph{Knights of the Dawn} and a culture of logistics.  
XP Examples: Enhance Self (stamina, logistics), Acquire Assets (ferry rights, keeps), Learn Talents (road-law, dawn-muster).  

\subsubsection{Vhasians}
Old nobility and fractious dukes. Value bloodlines, alliances, and grandeur.  
XP Examples: Enhance Self (Presence), Acquire Assets (estates, alliances), Learn Talents (courtly intrigue).  

\subsubsection{Ecktorians}
Imperial remnants, gladiatorial traditions, and Church of the Flame zeal.  
XP Examples: Enhance Self (Body/Presence), Acquire Assets (arena, church patronage), Learn Talents (oratory, zealotry).  

\subsubsection{Acasians}
Petty warlords and bravos. Known for Silkstrand and mercenary companies.  
XP Examples: Enhance Self (blade skill), Acquire Assets (bravo gangs, ships), Learn Talents (dueling flourishes).  

\subsubsection{Ubral}
Clannish highlanders and lowlanders; guerrilla fighters and dwarven allies.  
XP Examples: Enhance Self (Stealth, guerilla tactics), Acquire Assets (clanholds, blood oaths), Learn Talents (hit-and-run mastery).  

\subsubsection{Tulkani}
Wanderers, performers, and traders across the north and Amrathian basin.  
XP Examples: Enhance Self (performance, cunning), Acquire Assets (wagon caravans, hidden networks), Learn Talents (witchcraft, shadowbinding).  

---

\section{Dwarves (Aeler)}
Stone-founded, guild-bound, collectivist.  
Affinity: \emph{Stone-Sense}.  

\subsection{Sub-Types}
\subsubsection{Mountain Dwarves}
Deep kingdom dwellers; focused on giants, dragons, and underrealms. Prestige: Forge-Patriarchs.  

\subsubsection{Hill Dwarves}
Surface allies of humans; fight Ykrul and trolls. Prestige: Spirit Shield Warriors.  

\subsubsection{Edgewalkers}
Outcast adventurers, risk-takers, and opportunists abroad. 

\subsubsection{True Masons}
Itenerate stone-clerics, craftsmen who maintain Aeler infrastructure abroad.

\subsubsection{Spirit Shields}
Traditionalist, spiritual warriors who uphold the old ways.

---

\section{Elves (Lethai)}
Elves see themselves as the first people, divided by philosophy.  

\subsection{Sub-Types}
\subsubsection{Wood Elves (Lethai-al)}
“Mist People,” tribal and fey; cycle-focused. Prestige: Nature Rangers, Backlash Soothers.  

\subsubsection{High Elves (Lethai-thora)}
Sequestered exiles of the mind; memory and continuity. Prestige: Echo-Walkers.  

\subsubsection{Grass / Dusk / Sundered Elves}
Nomadic high elves estranged from both cousins.  

---

\section{Ykrul}
Pragmatic warbands of horse, boar, and warg.  
Affinity: \emph{Blood Memory}.  

\subsection{Sub-Types}
\subsubsection{Steppe Ykrul}
Pastoral nomads of the Violet Steppes.  

\subsubsection{Mountain Ykrul}
Stockier, fierce, more reclusive clans of the Aelerian highlands.  

\subsubsection{River Ykrul}
Wargboat raiders and coastal plunderers. Prestige: Warglords.  

---

\section{Gnomes (Aelinnel)}
Affinity: \emph{Numerical Insight}. Known for incomprehensible mathematics and craft.  

\subsection{Sub-Types}
\subsubsection{Rock Gnomes}
Burrow-dwellers of Wrentfell; gemstone engineers.  

\subsubsection{Forest Gnomes}
Valewood dwellers; more attuned to nature and fey.  

\subsubsection{Faetae (Winged Gnomes)}
Rare, fragile, and fae-blooded with insectile wings. Can fly but this often leads to complications.  

---

\section{Halflings (Aelaeram)}
Affinity: \emph{Hearth Sense}. Pastoral and resilient.  

\subsection{Sub-Types}
\subsubsection{Hill Halflings}
The most common; Amedell farmers and traders.  

\subsubsection{River Halflings}
Bargefolk and ferry-keepers along the Belworth.  

\subsubsection{Wander Halflings}
Caravaners and adventurers who leave the hearth to find fortune.   

\section{The XP Paths in Practice}
\section{Enhance Self}
Investing in raw power and skill.  
\begin{quote}
A Viterran knight spends XP to raise Body and Melee, becoming a duelist feared on the field.  
\end{quote}

\section{Acquire Assets}
Investing in the world around you.  
\begin{quote}
A Vhasian Duke invests XP into a Keep (off-screen) and house guards (on-screen).  
\end{quote}

\section{Learn Talents}
Investing in uniqueness.  
\begin{quote}
A Tulkani shadowbinder learns to weave concealment magic, turning night into an ally.  
\end{quote}

\section{Putting It Together}
Every player’s path blends the three. One may stand alone as a Solo, another weave allies as a Mastermind, another balance the two.  
Fate’s Edge rewards the choice, not just the outcome.  
The world reacts to where you put your weight—on the self, the world, or the myth you embody.

\section{Stock NPCs}

Not every companion, pet, or hireling needs to be custom-built.  
This section provides quick archetypes for \textbf{players} (on-screen allies) and \textbf{GMs} (encounters, foils, and rivals).

\section{Player Companions}
On-screen followers are bought with XP, and their abilities are measured by a \emph{Skill Cap}. They support but also create vulnerabilities.  

\subsection{Familiars \& Pets}
Bound by bond or training. Useful for scouting, flavor, and niche aid.  
\begin{description}[leftmargin=2cm]
  \item[Shadow-Cat] Cap 2 in Stealth. Can climb, slip through cracks, and carry small items. Backlash: draws bad omens.  
  \item[Crow Messenger] Cap 2 in Perception. Can deliver notes or warn of danger. Backlash: nosy, sometimes lies.  
  \item[Hound of the Fens] Cap 3 in Tracking. Keen nose, loyal defender. Backlash: loud bark alerts enemies.  
\end{description}

\subsection{Hirelings \& Followers}
Professionals who lend their skill to a player’s action.  
\begin{description}[leftmargin=2cm]
  \item[Bodyguard] Cap 4 in Melee. Grants bonus dice in combat, but draws fire.  
  \item[Scribe] Cap 3 in Lore. Keeps records, interprets contracts, whispers legal loopholes.  
  \item[Scout] Cap 3 in Survival. Knows paths, can extend a party’s range, but prone to wanderlust.  
\end{description}

\section{GM Stock NPCs}
These NPCs are not built with XP, but with intent. Each carries a role in pacing and tension.  

\subsection{Encounters}
\begin{description}[leftmargin=2cm]
  \item[Bandit Skirmisher] Body 2, Wits 2, Skills: Melee 2, Stealth 1. Light armor, opportunistic, breaks if pressed.  
  \item[Ykrul Rider] Body 4, Wits 3, Skills: Riding 3, Melee 3. Mobile, brutal, fights for plunder.  
  \item[Street Bravo] Presence 3, Body 2, Skills: Dueling 3. Quick to anger, duels to prove honor.  
\end{description}

\subsection{Foils \& Rivals}
\begin{description}[leftmargin=2cm]
  \item[Ambitious Scribe] Wits 3, Presence 3, Skills: Intrigue 3, Lore 2. Always knows a rumor, but hides motives.  
  \item[Mercenary Captain] Body 4, Spirit 3, Skills: Command 3, Melee 4. Pragmatic, can be ally or threat.  
  \item[Flame Preacher] Presence 4, Spirit 3, Skills: Oratory 4, Faith 3. Incites mobs, can spark holy trouble.  
\end{description}

\section{Gameplay Demo}

\section*{At the Table}
To illustrate the flow of play in \textit{Fate’s Edge}, here is a short example.  
The players are exploring an abandoned watchtower in the Mistlands.  

\subsection*{Scene: Climbing the Tower}

\textbf{GM:} The wooden stairs are rotted through. If you want to reach the beacon chamber, you’ll need to climb the stone wall.  

\textbf{Player (Rin, the Scout):} I’ll scramble up using my dagger for handholds. That’s Body + Athletics, right?  

\textbf{GM:} Yes. What’s your approach?  

\textbf{Rin:} Intricate — I describe how I wedge the dagger into the cracks, kicking off with precision.  

\textbf{GM:} Great. Roll 4 dice.  

\textit{The player rolls 6, 5, 1, 1. Two successes, two Complication Points. Because it’s Intricate, Rin rerolls both 1s: 7 and 2. Now three successes total.}  

\textbf{GM:} You make it to the top, but I’ll spend those two Complication Points.  
\textit{(Draws from the Deck of Consequences: Clubs 3 = “Equipment Strain.”)}  
Your dagger is stuck in the stone, bent and useless.  

\textbf{Rin:} Worth it — I pull myself onto the platform.  

---

\subsection*{Scene: The Beacon Chamber}

\textbf{GM:} The chamber reeks of smoke and salt. A hunched figure stirs the ashes — a Mistborn wraith.  

\textbf{Player (Kael, the Warlock):} I’ll cast a firebolt to drive it back. Channel first: Wits + Arcana.  

\textit{Kael rolls 5 dice: 8, 6, 3, 1, 1. Two successes, two Complication Points.}  

\textbf{Kael:} Detailed approach — I reroll one 1. That’s a 7. Three successes.  

\textbf{GM:} You’ve built strong Potential. Now Weave it with Fire.  

\textit{Kael rolls Wits + Fire Art: 6, 5, 2, 1. One success, one more Complication Point.}  

\textbf{GM:} The firebolt streaks out — the wraith hisses and withdraws. But I’ve got three Complication Points total.  
\textit{(Draws from Deck: Hearts 2 = “Emotional Cost.”)}  
The firebolt sparks panic in your companions — the Mistlands mist churns, and the shadows deepen. Everyone takes a level of Fatigue from the oppressive dread.  

---

\subsection*{Takeaway}
This example shows the rhythm of \textit{Fate’s Edge}:
\begin{itemize}
  \item Players describe their approach, choosing Basic, Detailed, or Intricate.  
  \item Dice are rolled; successes move the story forward, but Complication Points fuel the GM.  
  \item The GM uses Complications (or the Deck of Consequences) to keep tension alive.  
  \item Every roll advances the fiction, never just “succeed/fail.”  
\end{itemize}

\begin{tcolorbox}[title=Player Tip: How to Earn Boons,colback=black!2,colframe=black!40!white]
  Boons are not random handouts—they reward when you take \textbf{real risks} and \textbf{move the story forward}.  
  
  \begin{itemize}
    \item Take bold actions, even if they might fail.  
    \item Describe what you do in the fiction—make it vivid.  
    \item Accept complications and setbacks; they fuel your future.  
    \item Don’t “fish” with meaningless rolls—no story, no Boon.  
  \end{itemize}
  
  \textbf{Remember:} A Boon is proof that you dared, failed, and grew.
  \end{tcolorbox}

\section*{Min–Max Showcases (Player Options)}

\subsection*{The Unstoppable Blade (Solo Peak)}
\textbf{Concept:} A duelist honed to a lethal edge; controls any fight that involves their blade.

\paragraph{Dice Pools}
\begin{itemize}
  \item \textbf{Strike/Parry:} Body 4 + Melee 3 = \textbf{7d10}
  \item \textbf{Read Foe / Initiative:} Wits 3 + (no skill) = \textbf{3d10}
  \item \textbf{Social Baseline:} Presence 2 + (no skill) = \textbf{2d10}
\end{itemize}

\paragraph{XP Build (Target 32; rules-legal with deficit trades)}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Category} & \textbf{Pick} & \textbf{XP} \\
\midrule
Attributes & Body 4 & 12 \\
           & Wits 3 & 9 \\
Skills     & Melee 3 & 12 \\
Talents    & \emph{Battle Instincts} & 4 \\
\midrule
\multicolumn{2}{r}{\textit{Subtotal}} & \textbf{37} \\
\multicolumn{2}{r}{Starting Complications $\rightarrow$ XP (3 picks)} & \textbf{-3} \\
\multicolumn{2}{r}{Boon Bonds (4 Boons $\Rightarrow$ 2 XP)} & \textbf{-2} \\
\midrule
\multicolumn{2}{r}{\textit{Net Start}} & \textbf{32} \\
\bottomrule
\end{tabular}

\paragraph{Suggested Starting Complications (choose 3)}
\emph{Cursed Token}; \emph{Notoriety (dueling circles)}; \emph{Blood Debt}.  
\textbf{Boon Bonds:} Tie to two other PCs (mutual) for 4 Boons (2 XP).

\paragraph{Table Notes}
Always use \textbf{Intricate} descriptions in combat to re-roll 1s and add flourish. Terrifying in a fight, intentionally underpowered outside it — rely on allies for social/puzzle scenes.

\bigskip

\subsection*{The Ghost in the System (Mastermind Peak)}
\textbf{Concept:} A fixer who solves problems off-screen with networks, safehouses, and favors.

\paragraph{Dice Pools}
\begin{itemize}
  \item \textbf{Deception/Setup:} Presence 3 + Subterfuge 2 = \textbf{5d10}
  \item \textbf{Read Angles:} Wits 3 + Insight 2 = \textbf{5d10}
  \item \textbf{On-Screen Muscle:} (low; bring a bodyguard follower if desired)
\end{itemize}

\paragraph{XP Build (Target 32; rules-legal)}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Category} & \textbf{Pick} & \textbf{XP} \\
\midrule
Attributes & Presence 3 & 9 \\
           & Wits 3 & 9 \\
Skills     & Subterfuge 2 & 6 \\
           & Insight 2 & 6 \\
Off-Screen Asset & \textbf{Standard Safehouse Network} & 8 \\
\midrule
\multicolumn{2}{r}{\textit{Subtotal}} & \textbf{38} \\
\multicolumn{2}{r}{Starting Complications $\rightarrow$ XP (4 picks)} & \textbf{-4} \\
\multicolumn{2}{r}{Boon Bonds (4 Boons $\Rightarrow$ 2 XP)} & \textbf{-2} \\
\midrule
\multicolumn{2}{r}{\textit{Net Start}} & \textbf{32} \\
\bottomrule
\end{tabular}

\paragraph{Suggested Starting Complications (choose 4)}
\emph{Debt Mark}; \emph{Fractured Loyalty}; \emph{Blackmail}; \emph{Owed Favor}.  
\textbf{Boon Bonds:} Tie to two PCs for 4 Boons (2 XP).  
\textbf{Activation Fuel:} Use \textbf{1 Boon or 2 XP} to activate the Safehouse Network for a scene-level ask.

\paragraph{Table Notes}
You “win” by \emph{not rolling}. Spend Boons to open doors, farm Boons with controlled risks. Fragile if the Boon engine dries up.

\bigskip

\subsection*{The Voice of God (Face Peak)}
\textbf{Concept:} A presence that ends fights with words, ritual, and authority.

\paragraph{Dice Pools}
\begin{itemize}
  \item \textbf{Persuade/Compel:} Presence 3 + Diplomacy 3 + \emph{Silver Tongue} = \textbf{7d10}
  \item \textbf{Rally/Direct:} Presence 3 + Command 1 (+ talent if framed) = \textbf{4–5d10}
  \item \textbf{Resolve/Resist:} Spirit 3 = \textbf{3d10}
\end{itemize}

\paragraph{XP Build (Target 32; rules-legal with deficit trades)}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Category} & \textbf{Pick} & \textbf{XP} \\
\midrule
Attributes & Presence 3 & 9 \\
           & Spirit 3 & 9 \\
Skills     & Diplomacy 3 & 12 \\
           & Command 1 & 2 \\
Talents    & \emph{Silver Tongue} & 3 \\
\midrule
\multicolumn{2}{r}{\textit{Subtotal}} & \textbf{35} \\
\multicolumn{2}{r}{Starting Complications $\rightarrow$ XP (2 picks)} & \textbf{-2} \\
\multicolumn{2}{r}{Boon Bonds (2 Boons $\Rightarrow$ 1 XP)} & \textbf{-1} \\
\midrule
\multicolumn{2}{r}{\textit{Net Start}} & \textbf{32} \\
\bottomrule
\end{tabular}

\paragraph{Suggested Starting Complications (choose 2)}
\emph{Notoriety (public figure)}; \emph{Enemy at Large}.  
\textbf{Boon Bonds:} Tie to one PC for 2 Boons (1 XP), or two PCs for 4 Boons (2 XP) and bank 1 XP or add Insight~1.

\paragraph{Table Notes}
\emph{Intricate} speeches reroll 1s and add social flourishes (“he becomes a quiet informant”). You fold scenes without steel, but zealots/monsters resist your leverage. Keep allies between you and blades.

\bigskip

\paragraph{GM Sidebar (All Three)}
\begin{itemize}
  \item Reward \textbf{Intricate} declarations with rerolls and world-reactive flourishes.
  \item Pressure each archetype at its weakest axis (Blade: politics; Ghost: resource drought; Voice: brute force) while letting them shine often.
  \item Enforce Off-Screen Asset activation: \textbf{1 Boon or 2 XP} per concrete ask; if refused, show friction instead of hard no.
\end{itemize}
\begin{tcolorbox}[colback=gray!10!white,colframe=black,title=Absent PCs]
  Even if you miss a session, your character is still part of the world.  
  The group has two safe ways to use your presence:  
  
  \begin{itemize}
    \item \textbf{Off-Screen Help:} Your assets (safehouses, contacts, titles, etc.) can be tapped for preparation. This costs 2 Boons, or 1 Boon + 1 XP, paid by the party. Example: “Danny’s smuggler friend stashed a wagon.”  
    \item \textbf{Borrowed Action:} Once per session, with your consent, the GM may roll or act briefly as your PC in a scene. This always adds a new Complication to your sheet, scaled to the action. Example: “Danny invokes his title in court” → Complication: \emph{Overplayed Authority}.  
  \end{itemize}
  
  \textbf{Why?} This rule lets the story keep moving without you, while protecting your agency.  
  Your allies can call on you—but only sparingly, and it always leaves a mark.  
  \end{tcolorbox}

  \section{Archetype Showcase: Classic Heroes, Fate's Edge Style}

\noindent
Fate’s Edge does not provide rigid classes. Instead, it offers a toolkit of
attributes, skills, assets, and talents. To help players translate familiar
fantasy concepts into this system, here are ten archetypes inspired by
classic traditions, each reimagined to highlight \textbf{narrative weight,
off-screen assets, and consequences.}

\subsection{1. The Oath-Bound Blade (Paladin)}

\textbf{Classic Build:} A holy knight, beacon of faith and strength.\\
\textbf{Fate’s Edge Twist:} Power comes from an \textbf{Oath} so binding it
reshapes reality.\\
\textbf{XP Path:} High Spirit and Body. Skills in Command and Melee. Prestige
Ability: \emph{Oath-Keeper’s Word}.\\
\textbf{Twist:} Break the Oath and your powers risk becoming
\emph{Neglected}, requiring quests of atonement. Healing transfers harm to you,
and binding evil leaves scars.

\subsection{2. The Death-Speaker (Necromancer)}

\textbf{Classic Build:} Master of undead armies.\\
\textbf{Fate’s Edge Twist:} You bargain with the dead instead of commanding them.\\
\textbf{XP Path:} High Spirit and Wits. Skills in Arcana, Sway, Lore.\\
\textbf{Twist:} Summoning requires Boons or accepting Complications
(\emph{unfinished business, despair, rebellion}). An undead army is a
\emph{Major Asset} that can revolt if neglected.

\subsection{3. The Border-Warden (Ranger)}

\textbf{Classic Build:} Lone wilderness scout with beast companion.\\
\textbf{Fate’s Edge Twist:} A sworn custodian of a \emph{liminal territory}.\\
\textbf{XP Path:} High Wits and Survival. Assets: Lodge (Minor), Informants (Standard).\\
\textbf{Twist:} Power is limited geographically. Inside your domain, you gain
bonuses; outside, you are merely skilled. Every choice risks the balance of
your warded lands.

\subsection{4. The Guild-Approved Shadow (Rogue)}

\textbf{Classic Build:} A stealthy thief and trickster.\\
\textbf{Fate’s Edge Twist:} A licensed operative of a Guild or Noble’s Network.\\
\textbf{XP Path:} High Wits and Presence. Skills in Skulduggery, Stealth, Deception.\\
\textbf{Twist:} Calling on Guild resources costs Boons/XP. Failure
Compromises the Guild Asset, risking exile or disavowal. Theft becomes political leverage.

\subsection{5. The Spirit-Touched Outlander (Barbarian)}

\textbf{Classic Build:} Raging warrior of primal might.\\
\textbf{Fate’s Edge Twist:} Rage is trance-possession by an ancestral spirit.\\
\textbf{XP Path:} High Body, strong Spirit. Prestige Ability: \emph{Warglord}.\\
\textbf{Twist:} Gain +3 dice but risk Complications
(\emph{friend/foe confusion, compulsive pursuit}). Rage ends with Fatigue and isolation.

\subsection{6. The Scholar of Fractured Truths (Wizard)}

\textbf{Classic Build:} Spellbook-wielding intellectual.\\
\textbf{Fate’s Edge Twist:} Power comes from contradictory fragments of lore.\\
\textbf{XP Path:} High Wits and Spirit. Skills in Lore, Arcana, Research.\\
\textbf{Twist:} Each “spell” is a volatile truth. Activation requires Boons/XP,
and every spell risks a Complication. The Grimoire Asset may be
\emph{Compromised} by theft or misuse.

\subsection{7. The Caretaker of Cycles (Druid)}

\textbf{Classic Build:} Nature priest and shapeshifter.\\
\textbf{Fate’s Edge Twist:} A custodian of life-and-death balance.\\
\textbf{XP Path:} High Spirit, strong Survival and Command.\\
\textbf{Twist:} Healing demands an equal decay. Using druidic power creates
counterweight Complications. The Grove Asset reflects your successes or failures.

\subsection{8. The Chronicler of Consequences (Bard)}

\textbf{Classic Build:} Charismatic performer and inspirer.\\
\textbf{Fate’s Edge Twist:} A living archive whose songs define history itself.\\
\textbf{XP Path:} High Presence and Wits. Skills in Sway, Lore, Performance.\\
\textbf{Twist:} Inspiration weaves allies into a story. Failure twists it into tragedy,
producing Complications. Patronage and audiences are fragile Assets dependent on reputation.

\subsection{9. The Ascetic of the Unbound Body (Monk)}

\textbf{Classic Build:} Disciplined martial artist.\\
\textbf{Fate’s Edge Twist:} Detachment unbinds body and spirit.\\
\textbf{XP Path:} High Body and Spirit. Skills in Melee, Meditation, Lore.\\
\textbf{Twist:} Ignore fatigue or wounds by spending Boons, but harm
returns later as delayed Complications. Monastic Assets impose obligations and alienation.

\subsection{10. The Petitioner of a Silent God (Cleric)}

\textbf{Classic Build:} Priest channeling divine power.\\
\textbf{Fate’s Edge Twist:} Miracles are interpretations of divine silence.\\
\textbf{XP Path:} High Spirit and Presence. Skills in Diplomacy, Lore, Rituals.\\
\textbf{Twist:} Each divine act requires justification. If the logic falters, the GM twists
outcomes. The Temple Asset demands obedience, testing both faith and politics.

\bigskip

\noindent
These archetypes demonstrate that in Fate’s Edge, power is never free: every
ability, asset, and choice carries narrative cost. They are examples, not
restrictions, showing how classic fantasy concepts are reshaped by the
system’s focus on \textbf{consequence-driven storytelling.}

\begin{tcolorbox}[title=Fictional Snapshot: Captain Saikou Ira in Play,colback=gray!5,colframe=black!60!white]
The party infiltrates a merchant council session in Thepyrgos.
\begin{enumerate}
  \item \textbf{Observation:} Saikou rolls Wits + Insight (7d10). He notices a scribe hesitating every time a certain name is mentioned. A clue to divided loyalties.
  \item \textbf{Pressure:} With a quiet remark and steady gaze, he uses \emph{Psychological Pressure}. The scribe falters, allowing the party's diplomat to drive home the point with an extra edge.
  \item \textbf{Asset Play:} The group needs proof of smuggling routes. Saikou activates \emph{Naval Intelligence Contacts}. The GM describes a courier’s intercepted logbook, bought with favors and reputation.
  \item \textbf{Ritualist Moment:} Later that night, a cursed ledger exudes whispering voices. Saikou begins an exorcism using \emph{Ritual Casting}. Allies contribute dice, but a failed roll threatens to unleash a backlash. His \emph{Backlash Soothing} reduces the consequence, containing the voices with only a lingering fatigue.
  \item \textbf{Follower Spotlight:} When thugs confront the party outside, Sergeant Amari (Cap 4 Bodyguard) steps forward. Her +3 dice assist turns a desperate scuffle into a controlled rout.
\end{enumerate}
Throughout the scene, Saikou never swings a blade himself. His influence comes from foresight, pressure, and leveraging the right pieces on the board. His strength is that of a strategist: every move calculated, every risk deliberate, every victory costly but decisive.
\end{tcolorbox}

% ===== SRD: Asset-Linked Backgrounds =====
\section*{Asset-Linked Backgrounds}
\addcontentsline{toc}{section}{Asset-Linked Backgrounds}

\subsection*{What Backgrounds Are}
Backgrounds are optional, lore-forward packages that nudge play into vivid venues and give the GM natural places to spend Complication Points (CP). They do \emph{not} add clocks, stance edits, or extra actions.

\subsection*{Core Rule}
\begin{itemize}
  \item \textbf{Prerequisite Asset:} You must own the listed Asset (or acquire it now).
  \item \textbf{Talent (full cost):} Gain the specified Talent at its normal XP cost (recommended $\leq$ 4 XP).
  \item \textbf{Perk (pick exactly one; once per session):}
    \begin{enumerate}
      \item \textbf{+1 effect} on one opener in a keyed venue (ally or self).
      \item \textbf{Hazard $-1$} for one beat in a named environment (e.g., Deserts, Forests, Libraries, Waterways, Catacombs).
      \item \textbf{Mitigate CP} for the first on-theme twist in a specified scenario (reduce the first CP by 1).
      \item \textbf{Restore 1 follower} (clear 1 Exposure \emph{or} heal Harm 1).
      \item \textbf{Free upkeep} on one owned Asset.
    \end{enumerate}
  \item \textbf{Scene Keys (2):} Two venues where this background naturally appears (prompts only).
  \item \textbf{Hooks (GM CP sinks):} In keyed scenes, when the GM spends CP, choose \emph{one} listed Hook to fire.
  \item \textbf{Invite (optional):} A one-line behavior cue; if a player leans in, they may take 1 Boon (respect table Boon caps).
\end{itemize}

\subsection*{Guardrails}
\begin{itemize}
  \item Perks never change stance, create extra Initiative windows, pierce the global +3 Assist cap, or stack beyond normal tag rules (take the best one).
  \item Hooks are \emph{where} CP lands, not extra bookkeeping.
\end{itemize}

\subsection*{Hook Suit Markers (UI Language)}
Prefix each Hook with a suit symbol from the deck generator to suggest its flavor:
\quad \heartsuit~people/relationships \quad \clubsuit~systems/authority \quad \diamondsuit~resources/obligation \quad \spadesuit~danger/ominous.

% ===== Example Background (one; add more as needed) =====
\subsection*{Example: Vhasian Gentry}
\textbf{Requires Asset:} Vhasian Manor, Estate, or County\\
\textbf{Talent (full cost, 4 XP):} \emph{Influence at Court (Vhasian)} — +1 effect on formal petitions and protocol inside Vhasian courts and councils.\\
\textbf{Perk (pick 1; once/session):} \emph{Free upkeep} on one owned Asset.\\
\textbf{Scene Keys:} \emph{Court/Audience}, \emph{Noble Reception}.\\[2pt]
\textbf{Hooks (GM picks 1 on CP spend in keyed scenes):}
\begin{itemize}
  \item \clubsuit~A rival house “assists” by rerouting your petition to \emph{their} gatekeeper.
  \item \diamondsuit~Protocol demands a costly courtesy now (gift, sponsor, or public apology).
  \item \heartsuit~A jealous clerk bumps another case ahead; you owe someone face to keep momentum.
\end{itemize}
\textbf{Invite (optional):} “Defer to protocol before acting.”

% ===== End Section =====

% ===== SRD: Asset-Linked Backgrounds (Vililan Primer + Examples) =====
\section*{Asset-Linked Backgrounds — Vililan Primer}
\addcontentsline{toc}{section}{Asset-Linked Backgrounds — Vililan Primer}

\subsection*{Vililan: The Northern Continent (Lore Primer)}
North of the Amaranthine Sea, the continent of \textbf{Vililan} is a mosaic of post-imperial heirs, river-cities, fog coasts, and hard uplands. The old Utaran roads still bind the markets; the bells still mark law; the weather and water decide what endures.

\begin{description}
  \item[Ecktoria] Ledger and writ country: archives, toll-mints, and exam halls. Officials duel with precedent; petitioners duel with patience.
  \item[Acasia] The broken province: bridges, reed-roads, and a cosmopolitan port (Silkstrand) that keeps the north talking even when crowns don’t.
  \item[Vhasia] Hill-and-vale marches under old standards. Manor feuds and riding courts; proud levies that muster fast and argue faster.
  \item[Viterra] Fen and highland under a borrowed crown. Abbey-bells, causeways, and baronial law that remembers sovereignty.
  \item[Ubral] Iron-frontier culture on the eastern marches: garrisons, winter drills, and long memories for oaths kept and broken.
  \item[Theonan] Beacon coasts and cliff shrines; pilots and bell-keepers who read storm and shoal like scripture.
  \item[Vilikari] River syndics and convoy-law: wharf guilds, barge parliaments, and the art of moving goods without moving enemies.
  \item[Ostrikari] Ice-road runners and timber towns: sled relays, spring floods, and festivals that end in armistice instead of arrests.
  \item[Ykrul] Moor and barrow country: fog bells, hush rites, and wardens who know which doors should never be opened.
  \item[Aeler] Old high towers and fallen mysteries—anxious neighbors prefer their lore as flavor, not law. (Use sparingly unless your table wants deep archives.)
  \item[Linnic] Tower scholars and wind-hill observatories: star reckonings, drum-codes, and polite academic feuds with sharp edges.
  \item[Tulkani] Quarry towns and pit-runs: broken stone, brave crews, and contests that settle claims without lighting a war.
  \item[Kahfagian] Spur-steppes and caravan strings: horse-law, relay posts, and green-ribbon truces that make roads safer than palaces.
\end{description}

\paragraph{Backgrounds — Rules Recap}
Each Background is optional and requires a specific \textbf{Asset}. Pay the Talent at \textbf{full XP cost} (recommended $\leq$ 4 XP). Choose exactly \textbf{one Perk} (once/session): \emph{+1 effect (keyed venue opener)}, \emph{Hazard $-1$ (named environment) for one beat}, \emph{Mitigate first CP (scenario)}, \emph{Restore 1 follower (clear 1 Exposure or heal Harm 1)}, or \emph{Free upkeep on one owned Asset}. Hooks are where the GM spends CP in keyed scenes. Use suit markers as UI shorthand: \heartsuit~people/relationships \quad \clubsuit~systems/authority \quad \diamondsuit~resources/obligation \quad \spadesuit~danger/ominous.

% ===== Vhasian =====
\subsection*{Vhasian Gentry}
\textbf{Requires Asset:} Vhasian Manor, Estate, or County\\
\textbf{Talent (full cost, 4 XP):} \emph{Influence at Court (Vhasian)} — +1 effect on formal petitions and protocol inside Vhasian courts and councils.\\
\textbf{Perk (pick 1; once/session):} \emph{Free upkeep} on one owned Asset.\\
\textbf{Scene Keys:} \emph{Court/Audience}, \emph{Noble Reception}.\\[2pt]
\textbf{Hooks (GM picks 1 on CP spend in keyed scenes):}
\begin{itemize}
  \item \clubsuit~A rival house “assists” by rerouting your petition to \emph{their} gatekeeper.
  \item \diamondsuit~Protocol demands a costly courtesy now (gift, sponsor, or public apology).
  \item \heartsuit~A jealous clerk bumps another case ahead; you owe someone face to keep momentum.
\end{itemize}
\textbf{Invite (optional):} “Defer to protocol before acting.”

% ===== Ecktoria =====
\subsection*{Ecktorian Archivist}
\textbf{Requires Asset:} Archive Stacks, University Charter, or Scriptorium\\
\textbf{Talent (full cost, 3 XP):} \emph{Echo Reader} — hear space and lies; +1 effect when mapping by sound in stacks, vaults, or halls.\\
\textbf{Perk (pick 1; once/session):} \emph{Hazard $-1$} in \emph{Libraries/Archives} for one beat.\\
\textbf{Scene Keys:} \emph{Archive/Library}, \emph{Court of Inquiry}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \clubsuit~A citation challenge forces you to source a text \emph{now} before the hearing proceeds.
  \item \diamondsuit~A restricted ledger requires a donation, guarantor, or pledge.
  \item \spadesuit~A sealed sub-stack hums; opening it today will not stay quiet.
\end{itemize}
\textbf{Invite:} “Footnote your claim before you argue it.”

% ===== Acasia / Silkstrand =====
\subsection*{Silkstrand Span-Factor}
\textbf{Requires Asset:} Quay Office, Bridgehouse, or Processional Pennant\\
\textbf{Talent (full cost, 3 XP):} \emph{Processional Leader} — your first procession beat each session starts Controlled (per Talent).\\
\textbf{Perk (pick 1; once/session):} \emph{+1 effect} on one opener in a \emph{Bridge/Procession} scene.\\
\textbf{Scene Keys:} \emph{Bridge/Procession}, \emph{Festival Crowd}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~Toll-family feud flares; choosing a side brands you for the season.
  \item \diamondsuit~Porters expect a cut or favor to move the column.
  \item \spadesuit~A stray lantern wobbles; crowd mood tilts toward panic or zeal.
\end{itemize}
\textbf{Invite:} “Offer hospitality before demands.”

% ===== Viterra =====
\subsection*{Viterran Fenward Bailiff}
\textbf{Requires Asset:} Causeway Writ, Abbey Right, or Marsh Boat\\
\textbf{Talent (full cost, 4 XP):} \emph{Tide-Walker} — first water surge counts as Controlled for your side.\\
\textbf{Perk (pick 1; once/session):} \emph{Hazard $-1$} in \emph{Fens/Waterways} for one beat.\\
\textbf{Scene Keys:} \emph{Causeway Gate}, \emph{Fen Abbey}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \clubsuit~Abbey bell changes the order of cases mid-hearing.
  \item \diamondsuit~Tithe bales go missing; you inherit the shortfall until proven otherwise.
  \item \spadesuit~A spring surge arrives early; your call stands between pilgrims and flood.
\end{itemize}
\textbf{Invite:} “Name fair order even when it costs you.”

% ===== Ubral =====
\subsection*{Ubral Iron-Host Veteran}
\textbf{Requires Asset:} Garrison Quarters or Muster Charter\\
\textbf{Talent (full cost, 4 XP):} \emph{Rope Captain} — your calls steady teams; first bridge/assault beat starts Controlled.\\
\textbf{Perk (pick 1; once/session):} \emph{Restore 1 follower} (clear 1 Exposure or heal Harm 1 on a squadmate).\\
\textbf{Scene Keys:} \emph{Frontier Fort}, \emph{Recruiting Square}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~Old oath resurfaces; a survivor expects you to honor it.
  \item \clubsuit~Quartermaster’s book says your unit is “over strength” and trims rations.
  \item \spadesuit~Border flare-up: drills become live before you finish the briefing.
\end{itemize}
\textbf{Invite:} “Take the hard post without complaint.”

% ===== Theonan =====
\subsection*{Theonan Beacon-Keeper}
\textbf{Requires Asset:} Lighthouse/Beacon, Pilot Writ, or Bell Right\\
\textbf{Talent (full cost, 3 XP):} \emph{Lantern Ward} — nearby allies suffer reduced darkness penalties.\\
\textbf{Perk (pick 1; once/session):} \emph{Mitigate CP} in \emph{Storm/Coastal Night} scenes (reduce the first CP by 1).\\
\textbf{Scene Keys:} \emph{Coastal Tower}, \emph{Harbor in Storm}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \clubsuit~Harbor master demands you relight for a patron ship out of turn.
  \item \diamondsuit~Oil casks arrive watered; the supplier swears it wasn’t them.
  \item \spadesuit~A false beacon shows on the headland; someone is luring ships.
\end{itemize}
\textbf{Invite:} “Keep the light even if it blinds.”

% ===== Vilikari =====
\subsection*{Vilikari River Syndic}
\textbf{Requires Asset:} Wharf Syndicate Seat or Trade Charter\\
\textbf{Talent (full cost, 3 XP):} \emph{Ferry Whisperer} — +1 effect on quick deals at waterlines; pilots listen.\\
\textbf{Perk (pick 1; once/session):} \emph{+1 effect} on one opener in a \emph{Quay Market/Convoy} scene.\\
\textbf{Scene Keys:} \emph{Harbor Market}, \emph{River Convoy}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~Old rival factors your rate sheet to poach a client.
  \item \diamondsuit~Dock dues are “reinterpreted” mid-loading.
  \item \spadesuit~Rumors of press-gangs; crews vanish between bells.
\end{itemize}
\textbf{Invite:} “Cut a deal before you cut a line.”

% ===== Ostrikari =====
\subsection*{Ostrikari Ice-Road Runner}
\textbf{Requires Asset:} Sled Team, Ice-Road Writ, or Timber Relay\\
\textbf{Talent (full cost, 3 XP):} \emph{Bridge Runner} — ignore the first sway/slip penalty on spans and slick ways.\\
\textbf{Perk (pick 1; once/session):} \emph{Hazard $-1$} on \emph{Ice-Roads/Winter Crossings} for one beat.\\
\textbf{Scene Keys:} \emph{Frozen Ford}, \emph{Timber Relay}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \diamondsuit~Salt ration runs short; team health competes with speed.
  \item \clubsuit~Relay chief insists on the old route despite thaw warnings.
  \item \spadesuit~Hidden meltwater hollows the road ahead; the first to test pays.
\end{itemize}
\textbf{Invite:} “Trust the pole before the map.”

% ===== Ykrul =====
\subsection*{Ykrul Moor Warden}
\textbf{Requires Asset:} Fog Bell Right or Barrow Key\\
\textbf{Talent (full cost, 3 XP):} \emph{Night-Fire Tamer} — reduce the first fire-related Hazard tick each night scene.\\
\textbf{Perk (pick 1; once/session):} \emph{Mitigate CP} in \emph{Fog/Moor} scenes (reduce the first CP by 1).\\
\textbf{Scene Keys:} \emph{Barrow Field}, \emph{Moor Crossing}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \spadesuit~A will-o’-wisp answers the bell; someone follows the wrong light.
  \item \heartsuit~Mourners demand a rite you don’t recognize; offense is easy.
  \item \clubsuit~Ward posts were moved; jurisdiction is suddenly unclear.
\end{itemize}
\textbf{Invite:} “Name the hush before you name the danger.”

% ===== Linnic =====
\subsection*{Linnic Tower Scholar}
\textbf{Requires Asset:} Observatory Lease or Tower Right\\
\textbf{Talent (full cost, 3 XP):} \emph{Choir Seal} — lead a counting hymn to freeze a social rail for one beat.\\
\textbf{Perk (pick 1; once/session):} \emph{Mitigate CP} for \emph{Ritual/Scholarly} scenes (reduce the first CP by 1).\\
\textbf{Scene Keys:} \emph{Observatory}, \emph{Colloquium}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \clubsuit~A rival cites a conflicting ephemeris; the chamber splits.
  \item \diamondsuit~Your sponsor expects a conclusion that the stars don’t support.
  \item \heartsuit~A student’s blurted truth forces you to pick a side publicly.
\end{itemize}
\textbf{Invite:} “Count aloud; let the room breathe.”

% ===== Tulkani =====
\subsection*{Tulkani Pit-Runner}
\textbf{Requires Asset:} Quarry License or Arena Surety\\
\textbf{Talent (full cost, 3 XP):} \emph{Lantern Champion} — +1 effect on nonlethal duels and festival trials.\\
\textbf{Perk (pick 1; once/session):} \emph{+1 effect} on one opener in a \emph{Contest/Quarry} scene.\\
\textbf{Scene Keys:} \emph{Pit Run}, \emph{Stone Market}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~A cousin’s wager puts your honor on the line.
  \item \diamondsuit~Stone weights are swapped; a rival wants your lane disqualified.
  \item \spadesuit~A rope snaps mid-trial; it’s save or spectacle.
\end{itemize}
\textbf{Invite:} “Win clean when eyes watch.”

% ===== Kahfagian =====
\subsection*{Kahfagian Spur Captain}
\textbf{Requires Asset:} Spur Charter, Relay Post, or Herd Right\\
\textbf{Talent (full cost, 3 XP):} \emph{Safe-Conduct Bearer} — ignore the first checkpoint challenge once per journey under your pennant.\\
\textbf{Perk (pick 1; once/session):} \emph{Restore 1 follower} (clear 1 Exposure or heal Harm 1 on a beast-handler/escort).\\
\textbf{Scene Keys:} \emph{Steppe Relay}, \emph{Frontier Market}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \clubsuit~Relay steward downgrades your stamp; your escorts lose precedence.
  \item \diamondsuit~A tribute is due along the Green Ribbon and you are the collector—today.
  \item \spadesuit~Dust-ride signals ambush; your lane is the bait or the answer.
\end{itemize}
\textbf{Invite:} “Announce safe conduct like you mean it.”

% ===== SRD: Asset-Linked Backgrounds — Archetype Mappings (Aeler & Kin) =====
\section*{Asset-Linked Backgrounds — Archetype Mappings (Aeler \& Kin)}
\addcontentsline{toc}{section}{Asset-Linked Backgrounds — Archetype Mappings (Aeler \& Kin)}

\paragraph{Rules Recap (summary)}
Each Background is optional and requires a specific \textbf{Asset}. Pay the listed \textbf{Talent at full XP cost} (recommended $\leq$ 4 XP). Pick exactly \textbf{one Perk} (once/session): \emph{+1 effect (keyed venue opener)}, \emph{Hazard $-1$ (named environment) for one beat}, \emph{Mitigate first CP (scenario)}, \emph{Restore 1 follower (clear 1 Exposure or heal Harm 1)}, or \emph{Free upkeep (one owned Asset)}. Hooks are where the GM spends CP in keyed scenes. Prefix Hooks with suit markers: \heartsuit~people/relationships \quad \clubsuit~systems/authority \quad \diamondsuit~resources/obligation \quad \spadesuit~danger/ominous.

% ——— AELER: Wizard/Scholar archetype ———
\subsection*{Aeler Tower Savant}
\textbf{Requires Asset:} Aeler \emph{Sanctum Lease} or \emph{Grimoire Stacks}\\
\textbf{Talent (full cost, 3 XP):} \emph{Choir Seal} — lead a counting hymn to freeze a social rail for one beat.\\
\textbf{Perk (pick 1; once/session):} \emph{Mitigate CP} in \emph{Ritual/Arcane} scenes (reduce the first CP by 1).\\
\textbf{Scene Keys:} \emph{Sanctum/Tower}, \emph{Archive/Lecture}.\\[2pt]
\textbf{Hooks (GM picks 1 on CP spend in keyed scenes):}
\begin{itemize}
  \item \clubsuit~Ward custodian declares a breach of rite; seals must be audited before you proceed.
  \item \diamondsuit~Rare reagent is short; pay now or accept a weaker substitute with strings.
  \item \spadesuit~Dormant ward hums awake; backlash threatens your nearest ally.
\end{itemize}
\textbf{Invite:} “Name the rule before you bend it.”

% ——— LETHAI: Rogue/Veil archetype ———
\subsection*{Lethai Veil-Warden}
\textbf{Requires Asset:} \emph{Shadowhouse Key}, \emph{Veil Charter}, or \emph{Safehouse}\\
\textbf{Talent (full cost, 3 XP):} \emph{Mist-Step} — slip through fog lanes; +1 position when obscured.\\
\textbf{Perk (pick 1; once/session):} \emph{Hazard $-1$} in \emph{Shadows/Alleys/Night} for one beat.\\
\textbf{Scene Keys:} \emph{Night/Shadow}, \emph{Alley/Rooftop}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~An old contact recognizes your gait despite the mask; secrecy leaks.
  \item \clubsuit~Watch sergeant cites Veil restrictions here; your license is suddenly “conditional.”
  \item \spadesuit~A rival shadow lays a light-snare; the first misstep blinds rather than bruises.
\end{itemize}
\textbf{Invite:} “Leave no light where you pass.”

% ——— AELINNEL: Druid/Bard archetype ———
\subsection*{Aelinnel Grove Cantor}
\textbf{Requires Asset:} \emph{Grove Lease}, \emph{Sacred Oak Right}, or \emph{Shrine Glade}\\
\textbf{Talent (full cost, 4 XP):} \emph{Storm Courtier} — work clean in downpour; +1 effect on speeches in rain.\\
\textbf{Perk (pick 1; once/session):} \emph{Restore 1 follower} (clear 1 Exposure or heal Harm 1—beasts/companions included).\\
\textbf{Scene Keys:} \emph{Sacred Grove}, \emph{Forest March}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~Local clan expects a blessing before passage; refusing costs trust.
  \item \diamondsuit~Harvest tithe is due in the singing grove; pay in goods or service.
  \item \spadesuit~Root-ward reacts to iron; stepping wrong angers the glade.
\end{itemize}
\textbf{Invite:} “Sing first, then cut.”

% ——— AELAEREM: Artificer/Abjurer archetype ———
\subsection*{Aelaerem Wardwright}
\textbf{Requires Asset:} \emph{Wardstone Key}, \emph{Workshop}, or \emph{Forge Charter}\\
\textbf{Talent (full cost, 4 XP):} \emph{Bell Smith's Hand} — repair clappers and mounts in the field.\\
\textbf{Perk (pick 1; once/session):} \emph{Mitigate CP} on \emph{Trap/Ward} interactions (reduce the first CP by 1).\\
\textbf{Scene Keys:} \emph{Workshop/Forge}, \emph{Ruin/Barrier}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \clubsuit~Guild inspector demands a license stamp now or work stops.
  \item \diamondsuit~Patron swaps materials mid-job; performance guarantees still hold.
  \item \spadesuit~Old seal vents; the safe route becomes loud and time-limited.
\end{itemize}
\textbf{Invite:} “Measure twice, wake once.”

% ——— OATH/KNIGHT: Paladin archetype ———
\subsection*{Oath-Pennant Keeper}
\textbf{Requires Asset:} \emph{Oath-Pennant} or \emph{Chapter House}\\
\textbf{Talent (full cost, 3 XP):} \emph{Truce Broker} — invoke hospitality between rivals for one exchange.\\
\textbf{Perk (pick 1; once/session):} \emph{Mitigate CP} in \emph{Sanctuary/Vigil} scenes (reduce the first CP by 1).\\
\textbf{Scene Keys:} \emph{Shrine/Sanctuary}, \emph{Roadside Parley}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~A petitioner claims your protection while enemies close; denying it costs honor.
  \item \clubsuit~Magistrate contests your pennant’s authority on this ground.
  \item \spadesuit~An oathbreaker uses truce to approach; their price is a blade.
\end{itemize}
\textbf{Invite:} “Stand between names and knives.”

% ——— SEPULCHER: Necromancer/Cleric (lawful death) archetype ———
\subsection*{Sepulcher Adept}
\textbf{Requires Asset:} \emph{Crypt Ward} or \emph{Sub-Temple Custody}\\
\textbf{Talent (full cost, 3 XP):} \emph{Bell-Keeper} — once per session, read a bell to spot a coming surge or lie.\\
\textbf{Perk (pick 1; once/session):} \emph{Hazard $-1$} in \emph{Catacombs/Barrows} for one beat.\\
\textbf{Scene Keys:} \emph{Funeral Court}, \emph{Catacomb/Barrow}.\\[2pt]
\textbf{Hooks:}
\begin{itemize}
  \item \heartsuit~Mourners demand a rite you don’t recognize; offense is easy and loud.
  \item \clubsuit~Curia restricts the bone road today; paperwork or patron required.
  \item \spadesuit~A restless echo stirs and follows any light you carry.
\end{itemize}
\textbf{Invite:} “Name the hush before the hymn.”

% ===== SRD: Prestige Backgrounds (Tier III Assets & Epic Hooks) =====
\section*{Prestige Backgrounds (Tier III Assets \& Epic Hooks)}
\addcontentsline{toc}{section}{Prestige Backgrounds (Tier III Assets \& Epic Hooks)}

\subsection*{What Makes It “Prestige”}
Prestige Backgrounds are optional packages tied to \textbf{Tier III} Assets and \textbf{high-cost} Talents. They steer play toward flagship setpieces and give the GM a single, clearly-signposted \textbf{Epic Hook} to cash a large bank of CP \emph{without mitigation}. No stance edits, no extra clocks.

\paragraph{Rules (Prestige variant)}
\begin{itemize}
  \item \textbf{Prerequisite Asset:} You must own the listed \emph{Tier III} Asset (or acquire it now).
  \item \textbf{Talent (full cost):} Gain the specified Talent at \emph{its normal XP cost}, including Prestige-tier Talents (typically 6–10 XP).
  \item \textbf{Perk (pick exactly one; once/session):}
    \begin{enumerate}
      \item \textbf{+1 effect} on one opener in a keyed venue (ally or self), or
      \item \textbf{Hazard $-1$} for one beat in a named environment (e.g., Storm/Coast, Waterways, Catacombs), or
      \item \textbf{Mitigate CP} for a specific scenario (reduce the first CP by 1), or
      \item \textbf{Restore 1 follower} (clear 1 Exposure \emph{or} heal Harm 1), or
      \item \textbf{Free upkeep} on one owned Asset.
    \end{enumerate}
  \item \textbf{Scene Keys (2):} Two venues where this background naturally appears (prompts only).
  \item \textbf{Hooks (GM CP sinks):} In keyed scenes, when the GM spends CP, choose \emph{one} listed Hook to fire (use suit markers: \heartsuit people \quad \clubsuit systems \quad \diamondsuit resources \quad \spadesuit danger).
  \item \textbf{Epic Hook (GM CP dump):} In a keyed scene, the GM may declare the Epic Hook and \emph{spend any amount of banked CP}. This spend \textbf{cannot be reduced or prevented} by “Mitigate CP” perks, Network strings, or similar softeners. Players may still use normal resistance/soaks per core rules.
\end{itemize}

% ===== Example Prestige Background =====
\subsection*{Theonan Storm-Beacon Lord (Prestige)}
\textbf{Requires Asset (Tier III):} Grand Beacon (ocean headland) \emph{or} Harbor Light Consortium\\
\textbf{Talent (full cost, 8 XP):} \emph{Warden of Lanterns} — night scenes near you begin with darkness penalties reduced for allies.\\
\textbf{Perk (pick 1; once/session):}
\begin{itemize}
  \item \emph{Mitigate CP} in \emph{Storm/Coastal Night} scenes (reduce the first CP by 1), \emph{or}
  \item \emph{Hazard $-1$} in \emph{Waterways} for one beat, \emph{or}
  \item \emph{Free upkeep} on the Grand Beacon/Consortium for this session.
\end{itemize}
\textbf{Scene Keys:} \emph{Coastal Tower}, \emph{Harbor in Storm}.\\[2pt]
\textbf{Hooks (GM picks 1 on CP spend in keyed scenes):}
\begin{itemize}
  \item \clubsuit~Harbor master demands you \emph{relight out of turn} for a patron vessel; compliance angers pilots waiting on the bar.
  \item \diamondsuit~Oil casks arrive \emph{watered}; replacing them \emph{now} costs favors or coin you didn’t budget.
  \item \spadesuit~A \emph{false beacon} shows on the headland; someone is luring ships onto teeth.
\end{itemize}

\textbf{Epic Hook (unmitigable CP dump):} \spadesuit~\emph{Black Squall on the Bar}.\\
When the bar bell tolls thrice, gale bands cross the harbor mouth. The GM may immediately spend any amount of banked CP. For each \emph{1 CP} spent, choose one:
\begin{itemize}
  \item Advance \emph{two different rails} by +1 each (e.g., Hazard, Crowd, Escape).
  \item Snap a mooring or capsize a small craft; nearest ally makes an immediate save or marks Harm 1.
  \item Drown the quay in spray and smoke; on the next beat, all non-beacon signals are \emph{Obvious} and contested.
\end{itemize}
At \emph{5+ CP} spent at once, add: \emph{the sea wall overtops}; a wedge of the venue becomes \emph{briefly impassable}.\\
\emph{Note:} “Mitigate CP” perks, Network strings, and similar softeners do not apply to this spend. Standard resistance/soak still applies to resulting Harm or fallout.

\textbf{Invite (optional):} “Keep the light even if it blinds.”\\
\textbf{Tag:} \emph{Pilot-lord; storms hear your name.}

% ===== SRD: Prestige Backgrounds — Additional Entries =====

\subsection*{Aeler Forge-Master (Prestige)}
\textbf{Requires Asset (Tier III):} Great Forge–Citadel\\
\textbf{Talent (full cost, 18 XP):} \emph{Forge-Patriarch} — command great works; within your works you set cadence, safety, and yield for heavy craft and metalwork (per Talent text).\\
\textbf{Perk (pick 1; once/session):}
\begin{itemize}
  \item \emph{Hazard $-1$} in \emph{Foundry/Smelter} environments for one beat, \emph{or}
  \item \emph{Restore 1 follower} (clear 1 Exposure \emph{or} heal Harm 1 on a crew hand), \emph{or}
  \item \emph{Free upkeep} on the Forge–Citadel this session.
\end{itemize}
\textbf{Scene Keys:} \emph{Citadel Foundry Floor}, \emph{Guild Moot at the Pour}.\\[2pt]
\textbf{Hooks (GM picks 1 on CP spend in keyed scenes):}
\begin{itemize}
  \item \clubsuit~Guild tribune invokes Code–Red: a safety halt unless an inspector is appeased \emph{now}.
  \item \diamondsuit~Alloy shipment proves impure; pay to refine on the spot or accept brittle yield with strings.
  \item \heartsuit~Rival pattern-rights clan arrives with elders; proceeding without them costs standing.
\end{itemize}

\textbf{Epic Hook (unmitigable CP dump):} \spadesuit~\emph{Deep–Fire Surge}.\\
Stoked draft, slag pop, and a climbing chimney column threaten the whole work. The GM may immediately spend any amount of banked CP. For each \emph{1 CP} spent, choose one:
\begin{itemize}
  \item Advance \emph{two different rails} by +1 each among: \emph{Hazard} (heat, sparks), \emph{Crowd} (crew panic), \emph{Curfew} (city watch intervention), \emph{Sanctity} (guild code invoked).
  \item Drop a gantry or crack a crucible; the nearest PC/follower must save or mark \emph{Harm 1}; affected square becomes \emph{temporarily impassable}.
  \item Kill ventilation for a beat; on the next beat, \emph{all calls are Obvious} and verbal coordination falters.
\end{itemize}
At \emph{5+ CP} spent at once, add: \emph{chimney breach}—a section of venue is sealed by fire curtain or collapse until addressed on-screen.\\
\emph{Note:} “Mitigate CP” perks, Networks, and similar softeners do not reduce this spend. Standard resistance/soak still applies to resulting Harm or fallout.

\textbf{Invite (optional):} “Measure twice; wake the heat once.”\\
\textbf{Tag:} \emph{Master of the Pour; codes in the blood.}

% ——————————————————————————————————————————————

\subsection*{Oshiiran Ledgermaster (Prestige)}
\textbf{Requires Asset (Tier III):} Grand Exchange Charter\\
\textbf{Talent (full cost, 16 XP):} \emph{Master of Coin} — set precedence, unwind nets, and bind settlements across houses and ports (per Talent text).\\
\textbf{Perk (pick 1; once/session):}
\begin{itemize}
  \item \emph{+1 effect} on one \emph{negotiation/settlement} opener in a keyed venue, \emph{or}
  \item \emph{Mitigate first CP} in a \emph{Market/Exchange} scene (reduce the first CP by 1), \emph{or}
  \item \emph{Free upkeep} on the Grand Exchange Charter this session.
\end{itemize}
\textbf{Scene Keys:} \emph{Grand Exchange Hall}, \emph{Dockside Ledger–House}.\\[2pt]
\textbf{Hooks (GM picks 1 on CP spend in keyed scenes):}
\begin{itemize}
  \item \clubsuit~Customs board opens an emergency review; manifests are sequestered at a bad moment.
  \item \diamondsuit~Liquidity squeeze: a counterparty defaults; you must post bond or watch rights lapse.
  \item \heartsuit~A guild cousin arrives with a “family favor” that conflicts with today’s deal.
\end{itemize}

\textbf{Epic Hook (unmitigable CP dump):} \spadesuit~\emph{Audit Cascade}.\\
A cross-ledger variance fans outward—sureties wobble, credit freezes, rates spike. The GM may immediately spend any amount of banked CP. For each \emph{1 CP} spent, choose one:
\begin{itemize}
  \item Advance \emph{two different rails} by +1 each among: \emph{Crowd} (angry floor), \emph{Curfew} (enforcement clamp), \emph{Sanctity} (injunctions), \emph{Hunt} (tax agents, auditors), \emph{Escape} (blocked exits \& holds).
  \item \emph{Freeze one named Asset} (permit, bonded warehouse, ship share) for this scene; it cannot be activated until the scene turns or you placate an official on-screen.
  \item Trigger a \emph{margin call}: the acting PC must stake a Boon, favor, or public concession \emph{now} or lose face; refusal advances \emph{Crowd +1}.
\end{itemize}
At \emph{5+ CP} spent at once, add: \emph{Exchange halt}—all trade actions in the venue are treated as \emph{public and contested} until you restore order on-screen.\\
\emph{Note:} “Mitigate CP” perks, Networks, and similar softeners do not reduce this spend. Standard resistance/soak still applies to resulting Harm or fallout.

\textbf{Invite (optional):} “Open the books before the knives.”\\
\textbf{Tag:} \emph{Seal-bearer; markets move when you breathe.}

% ===== SRD: Comprehensive Talent Compendium for Fate's Edge =====
\section*{Comprehensive Talent Compendium}
\addcontentsline{toc}{section}{Comprehensive Talent Compendium}

\noindent
\textbf{Design guardrails (table-facing)} — Talents change \emph{position, effect, or rails}. They do not pierce the global \textbf{+3 Assist} ceiling, do not add extra actions (Initiative windows), and do not create Diamonds. “+1 die” in legacy text is translated here to \emph{start Controlled}, \emph{+1 effect}, \emph{Mitigate first CP}, or a targeted rail shift (\emph{Hunt/Escape/Hazard/Crowd/Curfew/Sanctity} $\pm1$) as appropriate.

\subsection*{Early Talents (3–5 XP)}
\begin{description}
  \item[\textbf{Battle Instincts} — 4 XP.] Once/scene, re-roll a failed \emph{defense} check. (Rerolls do not stack with other rerolls.)
  \item[\textbf{Silver Tongue} — 3 XP.] On a \emph{negotiation or fast-talk opener} in a keyed venue, gain \emph{+1 effect}. (Once/scene.)
  \item[\textbf{Stone-Sense} — 5 XP \emph{[Dwarf/Stone-trained]}] In contiguous stone, read flaws and load paths. Once/scene, grant \emph{start Controlled} to a move/breach using your read; additionally you may call \emph{Hazard $-1$} in a single 5–10 ft stone zone for one beat.
  \item[\textbf{Iron Stomach} — 3 XP.] Resist mundane toxins: the first toxin/rot \emph{Mitigate CP 1} this scene; on success, you ignore incidental nausea penalties for the scene.
  \item[\textbf{Versatile} — 3 XP \emph{[Human]}] Once/session, treat an untrained, adjacent task as if \emph{Trained} for position/effect (no dice added); or convert a “Limited” outcome to “Standard” on a clever repurpose.
  \item[\textbf{Guild Ties} — 5 XP.] Once/session, declare a \emph{Guild String}: immediately set an ally’s next on-venue action to \emph{start Controlled} \emph{or} \emph{Hunt/Escape $-1$}. (Creates a named social Hook for the GM.)
  \item[\textbf{Backlash Soothing} — 5 XP \emph{[Woodwise]}.] Once/scene in natural terrain, reduce magical \emph{Backlash} by \textbf{1 CP} \emph{or} convert \textbf{1 CP} to \emph{Fatigue 1} on the caster.
  \item[\textbf{Blood Memory} — 4 XP \emph{[Ykrul]}] After a violent scene, name a learned tactic; in the next scene against the same style, your first counter gains \emph{+1 effect}.
  \item[\textbf{Familiar Bond} — 9 XP.] Gain a spirit familiar (counts as an \emph{On-Screen follower}). No upkeep; cap \textbf{2} familiars (cap \textbf{3} with a prestige pick). Familiars assist with \emph{+Cap} in-role or \emph{+1} off-role on an \emph{intricate} description (total Assist still capped at +3).
  \item[\textbf{Processional Leader} — 3 XP.] Your first \emph{procession} beat each session \emph{starts Controlled}.
  \item[\textbf{Truce Broker} — 3 XP.] Declare hospitality for one exchange; no hostile actions without public cost.
  \item[\textbf{Lantern Ward} — 3 XP.] Reduce darkness penalties for nearby allies; once/scene, a night opener \emph{starts Controlled}.
  \item[\textbf{Route Whisper} — 2 XP.] On a declared route, your team’s first crossing gains \emph{+1 effect} \emph{or} \emph{Escape $-1$}. (Once/scene.)
  \item[\textbf{Hand Signals} — 2 XP.] When you set a silent plan, the next ally acting on it gains \emph{+1 effect}. (Once/beat.)
  \item[\textbf{Echo Mapping} — 2 XP.] (If taken as a Talent instead of Skill.) Read space by sound; grant an ally \emph{start Controlled} on one move/breach using your read (catacombs/archives).
  \item[\textbf{Bridge Runner} — 3 XP.] Ignore the first sway/slip penalty on spans and rigging each scene.
\end{description}

\subsection*{Mid-Tier Talents (6–10 XP)}
\begin{description}
  \item[\textbf{Commanding Presence} — 8 XP.] When an ally follows your explicit order, \emph{start Controlled} for that action (once/ally/scene).
  \item[\textbf{Warden of Lanterns} — 8 XP.] Night scenes near you begin with reduced darkness penalties; once/scene, name one ally’s night opener to \emph{start Controlled}.
  \item[\textbf{Storm-Caller} — 8 XP \emph{[Dhaharan/Sihai]}] Once/scene, influence local weather: choose \emph{Hazard $-1$ (wind/rain)} in one lane for a beat \emph{or} \emph{Mitigate CP 1} from storm surge.
  \item[\textbf{Beast-Tongue} — 8 XP \emph{[Ykrul]}] Parley with local beasts; once/scene, treat a suitable creature as a \emph{temporary Cap 2 follower} for one task.
  \item[\textbf{Rune-Keeper} — 8 XP \emph{[Aeler/Dwarf]}] Once/scene, scribe a quick ward: either \emph{Mitigate first CP 1} at this threshold or impose \emph{Hazard $-1$} vs intrusion for one beat.
  \item[\textbf{Blood-Ties Oath} — 8 XP \emph{[Fhara]}] When oath/honor is at stake, your first relevant action gains \emph{+1 effect} and a \emph{Protected Turn}.
  \item[\textbf{Guild Ledgercraft} — 8 XP \emph{[Oshiiran]}] Once/scene at market/exchange, pick one: \emph{start Controlled} for settlement \emph{or} \emph{Mitigate first CP 1} from a rate/inspection twist.
  \item[\textbf{Echo-Courtier} — 8 XP \emph{[Aeler]}] Once/scene in formal society, restage perception: shift one social rail \emph{$-1$} \emph{or} \emph{Mitigate first CP 1} from a “memory” twist (truth still matters).
  \item[\textbf{Battle Chant} — 8 XP \emph{[Ubral/Viterra]}] Allies within earshot ignore their \emph{first Fatigue 1} from exertion this combat.
  \item[\textbf{Whisper-Lord} — 8 XP \emph{[Sekogo/Oshiiran]}] Once/session, ask one pointed question about a named target’s \emph{current} status; the GM answers truthfully and adds a social Hook.
  \item[\textbf{River-Sovereign} — 8 XP \emph{[Dhaharan]}] Once/scene on water, call the right-of-way: \emph{Hazard $-1$} in your lane for a beat \emph{or} treat an ally’s maneuver as \emph{start Controlled}.
  \item[\textbf{Tide-Walker} — 8 XP.] The first surge in a water scene \emph{counts as Controlled} for your side; once/scene you may also \emph{Hazard $-1$} on a 10-ft square for one beat.
\end{description}

\subsection*{Leadership Talents (Scene Tempo)}
\begin{description}
  \item[\textbf{Coordinated Assault} — 8 XP.] When you Lead, you may direct \emph{up to two} helpers on the same roll. \emph{Total Assist still caps at +3.} If both are followers, it \emph{consumes this scene’s Follower Initiative}; each helper marks Exposure/Harm as fiction dictates.
  \item[\textbf{Master Coordinator} — 15 XP.] Once/scene after a successful Lead, choose: \emph{add +1 effect} to the team’s result \emph{or} \emph{open a second Initiative window} this scene. You take \emph{Fatigue 1}. (Does not raise the Assist cap.)
  \item[\textbf{Banner-Bearer} — 10 XP.] Once/battle, on a massed exchange, \emph{Mitigate the first 2 CP} from the group’s result (distributed as you narrate).
\end{description}

\subsection*{Prestige Talents (12+ XP)}
\begin{description}
  \item[\textbf{Echo-Walker} — 18–20 XP \emph{[Aeler]}] Once/arc, when you roll and generate \textbf{1–2 CP}, you may transmute that CP into an \emph{Omen}: bank \emph{+1 effect} on your next action this scene (cannot stack).
  \item[\textbf{Warglord} — 18 XP \emph{[Ykrul]}] Once/campaign, unify fractious bands under one banner for a single operation: treat disparate allies as a \emph{single Cap 3 unit} for one scene (narrative costs apply).
  \item[\textbf{Spirit-Shield} — 15 XP \emph{[Dwarf]}] Once/scene, on an ally’s result, \emph{reduce CP by up to 2}; you mark \emph{Fatigue 1} and accept \emph{Backlash 1} (narrative).
  \item[\textbf{Shadowbinder} — 16 XP \emph{[Tulkani/Veil]}] Once/scene in dim/shadow, treat the \emph{first detection} vs you as \emph{Controlled} and set \emph{Hunt $-1$}. (Not invisibility; leaves traces.)
  \item[\textbf{Forge-Patriarch} — 18 XP \emph{[Aeler/Dwarf]}] Within a Great Forge-Citadel, set cadence: one heavy craft sequence per scene \emph{starts Controlled}; you may \emph{Hazard $-1$} on the floor for one beat.
  \item[\textbf{Master of Coin} — 16 XP.] When you preside at an exchange, settlements and unwinds gain \emph{+1 effect}. Once/session in an Exchange scene, \emph{Mitigate first CP 1} from a rate/liquidity twist.
  \item[\textbf{Warden of Storms} — 18 XP \emph{[Sihai/Dhaharan]}] Once/arc, anchor or disperse a major storm. For this scene, your side’s first two \emph{storm} Hazards each suffer \emph{$-1$}.
  \item[\textbf{Mask-Bearer} — 16–17 XP \emph{[Nihon/Masked Courts]}] Once/scene while masked, you may \emph{escape recognition} at scene end if fiction allows; if used mid-scene, you gain a \emph{Protected Turn} but your presence is \emph{Obvious}.
  \item[\textbf{Star-Reader} — 18 XP \emph{[Linnic/Sihai]}] Once/arc under a predicted sky, one \emph{major} action gains \emph{start Controlled} and \emph{+1 effect}.
  \item[\textbf{Everflame Keeper} — 16 XP \emph{[Ecktoria]}] Once/scene, sanctify/purify: treat a corruption Hazard as \emph{Hazard $-1$} and grant a \emph{Protected Turn} for rite completion.
  \item[\textbf{Resonant Architect} — 18 XP \emph{[Dwarf]}] Once/arc, animate a place’s will: for one scene, treat the venue as a \emph{Cap 3 follower} you can direct (rails bite if abused).
  \item[\textbf{Sandseer} — 16–18 XP \emph{[Fhara]}] Once/scene in deserts, when an environmental threat would tick \emph{Hazard}, cancel that single tick \emph{or} convert it into \emph{+1 effect} on your navigation action.
\end{description}

\subsection*{Cultural Skill Paths (for Build Flavor)}
\begin{description}
  \item[\textbf{Scholarly Arts.}] \emph{Lore} (history/theology/occult), \emph{Research} (archives), \emph{Arcana} (theory/wards), \emph{Craft} (forge/alchemy/build), \emph{Ritual} (circle/shared casting).
  \item[\textbf{Social Arts.}] \emph{Diplomacy} (mediation), \emph{Sway} (street bargains), \emph{Performance} (oratory/mask), \emph{Subterfuge} (networks), \emph{Insight} (reads/lies).
  \item[\textbf{Martial Arts.}] \emph{Melee}, \emph{Ranged}, \emph{Brawl}, \emph{Athletics}, \emph{Tactics}.
  \item[\textbf{Wilderness Arts.}] \emph{Stealth}, \emph{Survival}, \emph{Navigation}, \emph{Animal Handling}, \emph{Scouting}.
\end{description}

\subsection*{Magic Specializations (Scene Levers)}
\begin{description}
  \item[\textbf{Elemental Focus} — 8 XP.] Pick an element (Fire/Water/Earth/Air). Once/scene with that Art, choose \emph{+1 effect} \emph{or} \emph{Mitigate Backlash 1} on your result.
  \item[\textbf{Backlash Resistant} — 8–10 XP.] The \emph{first} Backlash you suffer each scene is reduced by \textbf{1 CP} (min 0).
  \item[\textbf{Ritual Master} — 12 XP.] Lead multi-participant rituals without escalating Backlash tier; once/ritual you may convert \textbf{1 CP} to \emph{Fatigue 1}.
  \item[\textbf{Spell Anchor} — 8 XP.] Once/session, stabilize a failing spell by taking \emph{Fatigue 1}; the spell’s outcome stays at current stance/effect.
  \item[\textbf{Conduit Walker} — 15 XP.] Once/arc, open a safe conduit to Otherwhere: ask a single, bounded question or call a subtle boon (GM answers truthfully; may add an omen Hook).
\end{description}

\subsection*{Notes on Costs \& Cadence}
\begin{itemize}
  \item Costs reflect \emph{scene impact}, not simulation weight. If an entry feels table-warping, raise to the next band or add a fiction gate (lantern lit, mask on, hands free).
  \item “Once/scene” and “once/session/arc” are hard cadence caps; they \emph{replace} any looser language from legacy text.
  \item If a legacy talent says “+1 die,” use the nearest conversion: \emph{start Controlled}, \emph{+1 effect}, \emph{Mitigate first CP}, or \emph{rail $\pm1$}. Do not exceed the +3 Assist ceiling.
\end{itemize}

\section*{Comprehensive Asset Compendium}
\addcontentsline{toc}{section}{Comprehensive Asset Compendium}

\paragraph{Table-facing guardrails}
Assets are \emph{scene shapers}, not dice pumps. Any “+1 die” from legacy text is expressed as: \textbf{start Controlled}, \textbf{+1 effect}, \textbf{Mitigate first CP (by 1)}, or a named \textbf{rail} adjustment (\emph{Hunt/Escape/Hazard/Crowd/Curfew/Sanctity} $\pm 1$). Unless noted, each Off-Screen effect is \textbf{once per session}. No extra Initiative windows, no Diamond creation, no stacking identical tags (take the best one).

\subsection*{Minor Assets (4 XP)}
\paragraph{Signature Weapon.} Named heirloom or masterpiece; if lost/damaged, recovery is a story hook.
\emph{Off-Screen:} When this weapon would be broken/Compromised, \emph{Mitigate gear-break CP 1}. If still Compromised, you may declare a personal recovery scene this downtime to clear it.

\paragraph{Scholar’s Cell.} Quiet room with texts and desk.
\emph{Off-Screen:} One research opener in a keyed venue \emph{starts Controlled} \emph{or} takes \emph{+1 effect}.

\paragraph{Safehouse Network.} Hidden bolt-holes in one city.
\emph{Off-Screen:} Clear \emph{Hunt $-1$} between scenes \emph{or} begin the next on-site scene \emph{start Controlled} for entry.

\paragraph{Artisan’s Workshop.} Tools + bench for field-worthy repairs.
\emph{Off-Screen:} Clear \emph{Compromised} from one item.

\paragraph{Courier Network.} Trusty runners and dead-drops.
\emph{Off-Screen:} Deliver a message \emph{between scenes}; your next parley with that party \emph{starts Controlled} \emph{or} \emph{Mitigate first CP 1} on that negotiation.

\paragraph{Small Farm/Orchard.} Your larder and a little surplus.
\emph{Off-Screen:} \emph{Provision}: waive one minor Supply complication this session \emph{or} refresh one “consumable” tag on gear.

\paragraph{Local Tavern/Inn.} Neighborhood hub and rumor mill.
\emph{Off-Screen:} Your first local social opener \emph{starts Controlled}.

\paragraph{Minor Shrine/Temple.} Hearth for the spirit.
\emph{Off-Screen:} Clear \emph{Fatigue 1} \emph{or} \emph{Mitigate first morale CP 1} this session.

\paragraph{Fishing Boat/Barge.} River/harbor skiff with a pilot.
\emph{Off-Screen:} Treat one water crossing as \emph{start Controlled} \emph{or} ignore the first travel Supply loss.

\paragraph{Herbal Garden.} Beds of restorative greens.
\emph{Off-Screen:} Remove \emph{Fatigue 1} from one character.

\subsection*{Standard Assets (8 XP)}
\paragraph{Noble Title/Charter.} Recognized rank and privileges.
\emph{Off-Screen:} In authority venues, secure an audience; your first petition \emph{starts Controlled} \emph{or} you gain a \emph{Protected Turn}.

\paragraph{Guild Membership.} Standing in a working order.
\emph{Off-Screen:} A procurement/commission action \emph{starts Controlled} \emph{or} \emph{Mitigate first CP 1} from shortages/inspection.

\paragraph{Merchant Stall/Front.} A fixed face in the market.
\emph{Off-Screen:} \emph{Free upkeep} on one Minor Asset \emph{or} give \emph{+1 effect} on one barter/opener at your stall.

\paragraph{Spy Ring/Informants.} Eyes and ears on retainer.
\emph{Off-Screen:} A targeted inquiry \emph{starts Controlled} \emph{or} gains \emph{+1 effect}; the GM may add a named social Hook.

\paragraph{Mercenary Contract.} Reach for hired steel.
\emph{Off-Screen:} Introduce a temporary Cap 2 unit for one beat on-screen (consumes the scene’s Follower Initiative).

\paragraph{Trading Charter.} Right of trade along named routes.
\emph{Off-Screen:} At customs/gates, \emph{Mitigate first Diamonds/Clubs CP 1} \emph{or} your crossing opener \emph{starts Controlled}.

\paragraph{Library Archive.} Stacks and a helpful custodian.
\emph{Off-Screen:} For one deep topic, your research opener \emph{starts Controlled} and takes \emph{+1 effect} (scholarly venues).

\paragraph{Craftsman’s Quarter.} Shop + apprentices.
\emph{Off-Screen:} Produce one quality piece: clear \emph{Compromised} \emph{or} grant that item \emph{+1 effect} once this session.

\paragraph{Temple Complex.} Choirs, courts, and care.
\emph{Off-Screen:} Among the faithful, gain a \emph{Protected Turn} \emph{or} \emph{Mitigate first Sanctity/Crowd CP 1}.

\paragraph{Port Authority.} Berths, booms, and bells.
\emph{Off-Screen:} Priority handling: \emph{Mitigate first Curfew/Crowd CP 1} at docks \emph{or} treat loading/unloading as \emph{start Controlled}.

\subsection*{Major Assets (12 XP)}
\paragraph{City District License.} You run a ward’s levers.
\emph{Off-Screen:} In-district, one civic opener \emph{starts Controlled} \emph{or} set a local rail \emph{$-1$} (Crowd/Curfew) for a beat.

\paragraph{Regional Monopoly.} Exclusive right in a staple.
\emph{Off-Screen:} In market scenes, cancel the first \emph{Diamonds} CP this session.

\paragraph{Fortress/Stronghold.} A hard place held by your banner.
\emph{Off-Screen:} In siege/raid contexts, ignore the first \emph{Spades} (danger) complication this session.

\paragraph{Large Estate/Demesne.} Fields, mills, crews.
\emph{Off-Screen:} You do not suffer minor Supply complications this session \emph{or} grant \emph{free upkeep} to one additional Asset.

\paragraph{Naval Fleet Share.} Oars, sails, signal flags.
\emph{Off-Screen:} Ensure safe passage: one maritime travel segment \emph{starts Controlled} and \emph{Mitigate first Hazard CP 1}.

\paragraph{University College.} A faculty, a quad, a seal.
\emph{Off-Screen:} During downtime, one PC reduces the XP cost of a single \emph{Skill} advance by 1 (min 1) in a listed curriculum.

\paragraph{Banking Charter.} Letters, ledgers, liens.
\emph{Off-Screen:} In finance scenes, \emph{Mitigate first Diamonds CP 1}; if no Diamonds twist occurs, your settlement gains \emph{+1 effect}.

\paragraph{Diplomatic Enclave.} Flag, immunity, ink.
\emph{Off-Screen:} In foreign courts, ignore the first \emph{Clubs} (jurisdiction) complication \emph{or} your parley \emph{starts Controlled}.

\paragraph{Mine/Quarry Operation.} Raw stone, hard people.
\emph{Off-Screen:} Provide materials: clear \emph{Compromised} on a structure/gear \emph{or} grant \emph{+1 effect} to one build/breach.

\paragraph{Market Control.} You set the bell.
\emph{Off-Screen:} Declare \emph{Surplus/Scarcity} for one staple this session; your negotiation opener about that staple \emph{starts Controlled}.

\subsection*{Tier III Assets (16–20 XP)}
\paragraph{Grand Forge–Citadel.} City-sized heat and hammer.
\emph{Off-Screen:} Produce masterworks; ignore the first gear-related complication each session \emph{or} clear \emph{Compromised} across a batch.

\paragraph{Harbor Light Consortium.} Beacons from cape to cape.
\emph{Off-Screen:} Dictate right-of-way: an allied vessel’s maneuver \emph{starts Controlled} \emph{or} set a 10-ft water square to \emph{Hazard $-1$} for a beat.

\paragraph{Grand Exchange Charter.} Markets move when you breathe.
\emph{Off-Screen:} In exchange halls, cancel \emph{one Diamonds CP} this session; if already clean, take \emph{+1 effect} on a settlement/unwind.

\paragraph{Thepyrgosi Great Library.} The north’s memory palace.
\emph{Off-Screen:} For a major inquiry, research \emph{starts Controlled} and gains \emph{+1 effect}; also \emph{Mitigate first scholarly CP 1}.

\paragraph{Kahfagian Fleet Command.} Signals across the sea lanes.
\emph{Off-Screen:} Declare \emph{blockade or corridor}: impose \emph{Escape +1} on hostile sea exits \emph{or} ensure friendly passage \emph{start Controlled}.

\paragraph{Oshiiran Hydraulic Empire.} Water is wealth.
\emph{Off-Screen:} Set the \emph{Irrigated} tag for one route; cancel the first Supply/shortage \emph{Diamonds} CP this session.

\paragraph{Ashaani Palace Complex.} Keys to office and omen.
\emph{Off-Screen:} One political opener \emph{starts Controlled} and gains a \emph{Protected Turn}.

\paragraph{Valewood Grove Sanctuary.} Old roots, older vows.
\emph{Off-Screen:} In wilderness scenes, cancel one \emph{natural Hazard} tick \emph{or} \emph{Mitigate first CP 1} from weather/terrain.

\paragraph{Dwarven Mountain Hall.} Deep law, deeper craft.
\emph{Off-Screen:} Engineering/crafting opener \emph{starts Controlled} and gains \emph{+1 effect}.

\paragraph{Ykrul Steppe Confederation.} Ten thousand hooves.
\emph{Off-Screen:} On the steppe, treat one group movement as \emph{start Controlled} and apply \emph{Escape $-1$} for that leg.

\subsection*{Specialized Assets}
\paragraph{Magical Laboratory (10 XP).} Arcane benches, wards, and fumehoods.
\emph{Off-Screen:} In rituals here, reduce Backlash by \textbf{1 CP} once; a lab experiment opener \emph{starts Controlled}.

\paragraph{Healing Sanctuary (8 XP).} Beds, herbs, bells.
\emph{Off-Screen:} Treat a patient: clear \emph{Harm 1} or \emph{Fatigue 2} (downtime only).

\paragraph{Performance Venue (6 XP).} Stage + crowd.
\emph{Off-Screen:} Build reputation; your first \emph{Performance} opener \emph{starts Controlled} \emph{or} gains \emph{+1 effect}.

\paragraph{Observatory/Star Tower (8 XP).} Ephemerides and lenses.
\emph{Off-Screen:} Predict timing; one navigation/timing opener \emph{starts Controlled}.

\paragraph{Alchemical Garden (10 XP).} Beds of rare reagents.
\emph{Off-Screen:} Produce a potent draught; once this session, \emph{Mitigate first toxin CP 1} \emph{or} grant a one-use \emph{+1 effect} salve.

\paragraph{Training Academy (12 XP).} Masters, drills, and mats.
\emph{Off-Screen:} During downtime, \emph{reduce the XP cost of one Skill advance by 1} (min 1) for one trainee.

\paragraph{Intelligence Bureau (10 XP).} Analysts and cut-outs.
\emph{Off-Screen:} Avoid surprise this session; your first recon opener \emph{starts Controlled}.

\paragraph{Hospitality House (6 XP).} Beds, bread, and rumor.
\emph{Off-Screen:} Gain a \emph{Protected Turn} in a hosted parley \emph{or} \emph{Mitigate first Crowd CP 1}.

\paragraph{Storage Vault (6 XP).} Locks, ledgers, layers.
\emph{Off-Screen:} Ignore the first theft/sabotage complication targeting stored goods.

\paragraph{Communication Relay (8 XP).} Fires, flags, drums.
\emph{Off-Screen:} Coordinate afar: your next timed coordination \emph{starts Controlled} \emph{or} you \emph{Mitigate first Curfew CP 1} on response time.

\subsection*{Asset Upgrades \& Variants}
\paragraph{Enhanced Security (+2 XP).} Hardened doors, better watchers.
\emph{Effect:} The first theft/sabotage complication targeting this Asset each session is \emph{Mitigated (CP $-1$)}.

\paragraph{Expanded Scope (+3 XP).} Wider footprint or remit.
\emph{Effect:} Use this Asset’s Off-Screen effect \emph{twice per session} \emph{or} apply it across two adjacent venues (GM approval).

\paragraph{Magical Enhancement (+4 XP).} Warded, sanctified, or ensouled.
\emph{Effect:} Add one of: \emph{Mitigate Backlash 1 (once/session)} for rituals held here; \emph{Hazard $-1$} aura in a 10-ft zone for one beat (once/scene).

\paragraph{Political Connections (+3 XP).} Papers, patrons, precedence.
\emph{Effect:} In civic/authority scenes at this Asset, \emph{Mitigate first Clubs CP 1}.

\paragraph{Integration tip (optional).} If you’re using \emph{Asset-Linked Backgrounds}, note which Backgrounds key off the same venue; the GM can spend CP using those Background Hooks when this Asset is in play (no extra tracking needed).

% ===== End Compendium =====
\end{document}
