
\documentclass[11pt]{article}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{enumitem}
\usepackage{tcolorbox}
\usepackage{titlesec}
\usepackage{lmodern}

\titleformat{\section}{\normalfont\Large\bfseries}{}{0pt}{}
\titleformat{\subsection}{\normalfont\bfseries}{}{0pt}{}

\title{So My Dumb Friends Are Making Me Play This Stupid Game}
\date{}
\begin{document}
\maketitle

\section*{Step 1: Calm Down}
Yes, your friends roped you into another dice-rolling storytelling thing. Don’t panic. This one’s simple. You won’t need calculus, flowcharts, or to read 400 pages of fake history.  

\section*{Step 2: Make a Guy}
Your “character” only needs:  
\begin{itemize}[noitemsep]
    \item \textbf{Four Stats:} Body, Wits, Spirit, Presence (just pick 3, 2, 2, 1 in any order).  
    \item \textbf{Some Skills:} Melee, Diplomacy, Subterfuge, Lore, etc. Just pick 2 at level 2, 2 at level 1. Done.  
    \item \textbf{A Quirk:} One weird detail so the others don’t forget who you are (e.g., “Always carries a broom,” “Talks in rhyme”).  
\end{itemize}

\section*{Step 3: Roll Dice, Cause Problems}
When you do stuff:  
\begin{enumerate}[noitemsep]
    \item Add Stat + Skill = number of d10s to roll.  
    \item Each 6+ is a success. Each 1 is a “complication” (i.e., GM makes your life worse).  
    \item Describe what you’re doing. The fancier the description, the more you get to reroll 1s.  
\end{enumerate}

\section*{Step 4: Boons (The Fail Juice)}
Failing earns you \textbf{Boons}, which are basically “pity points.”  
\begin{itemize}[noitemsep]
    \item Spend 1 Boon to reroll a die.  
    \item Cash in 2 Boons between sessions for 1 XP (character growth).  
\end{itemize}
So, failing is actually good. You can’t lose, only \emph{lose stylishly}.  

\section*{Step 5: XP (The Bribe System)}
XP is the currency of life. You can:  
\begin{itemize}[noitemsep]
    \item Raise Stats or Skills (get better at rolling dice).  
    \item Buy Talents (special tricks).  
    \item Buy Assets (a tavern, a spy, a pet ferret, whatever).  
\end{itemize}
If you’re lazy: hoard XP, pretend you’ll spend it later.  

\section*{Step 6: Don’t Be a Jerk}
The only real rule is: be present in the scene. You can play “the mastermind with minions,” but you still have to show up. No off-screen overlords. No “my butler does it.”  

\section*{Step 7: Winning}
You don’t. It’s a story. But you can:  
\begin{itemize}[noitemsep]
    \item Be the Cool Hero.  
    \item Be the Dumb Disaster.  
    \item Accidentally marry an elf-queen.  
\end{itemize}
The real win is annoying your friends with how awesome/stupid your character becomes.  

\section*{Quick Table (So You Don’t Have to Ask)}
\begin{tcolorbox}[colback=gray!10, colframe=black, title=Cheat Sheet]
\begin{itemize}[noitemsep]
    \item Roll d10s: 6+ = success, 1 = complication.  
    \item Fail? Gain a Boon.  
    \item 1 Boon = reroll a die. 2 Boons = 1 XP.  
    \item XP buys stats, skills, talents, or assets.  
    \item Describe your action:  
    \begin{itemize}[noitemsep]
        \item Basic = roll straight.  
        \item Detailed = reroll 1 die that came up 1.  
        \item Intricate = reroll all 1s + add a flourish.  
    \end{itemize}
\end{itemize}
\end{tcolorbox}

\bigskip
\begin{center}
\textit{Congratulations. You’re now “playing.” You may go back to heckling your friends.}
\end{center}

\end{document}
