Looking at the SRD, I need to check the upkeep rules. From what I can see in the provided text, there's a discrepancy in how upkeep is calculated. Let me revise this section to match the SRD:

\chapter{Assets and Followers}
\label{ch:assets-followers}

Your character's influence extends beyond personal capabilities through \textbf{Assets} and \textbf{Followers}. These represent worldly possessions, connections, and allies that can solve problems, provide assistance, and shape the narrative.

\section{Understanding Assets and Followers}
\index{assets}\index{followers}

\subsection*{Key Differences}
\begin{itemize}
\item \textbf{Assets}: Off-screen resources that solve problems between scenes.
\item \textbf{Followers}: On-screen allies who assist during gameplay.
\item \textbf{Assets} change the fictional situation before you arrive.
\item \textbf{Followers} act alongside you in the moment.
\end{itemize}

\subsection*{Management Requirements}
Both require maintenance and carry risks:
\begin{itemize}
\item Regular upkeep costs (XP or downtime).
\item Vulnerability to complications and attacks.
\item Narrative consequences for misuse or neglect.
\end{itemize}

\section{Assets System}
\index{assets!system}

Assets are possessions, properties, or resources you control.

\subsection*{Asset Types and Costs}
\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Type} & \textbf{XP Cost} & \textbf{Establishment Time} \\
\midrule
Minor   & 4 XP  & 1 day \\
Standard& 8 XP  & 1 week \\
Major   & 12 XP & 1 month \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Asset Examples}
\begin{description}
\item[Minor Assets] Small shop, safehouse, minor title, basic workshop.
\item[Standard Assets] Noble title, guild membership, trading post, spy network.
\item[Major Assets] Fortress, city license, major enterprise, regional influence.
\end{description}

\subsection*{Using Assets}
Assets provide benefits in different ways:

\paragraph{Free Off-Screen Use}
Each asset has a specific off-screen effect you can use once per session:
\begin{itemize}
\item \textbf{Safehouse}: Provide secure lodging for the party.
\item \textbf{Spy Network}: Gather basic intelligence about a location.
\item \textbf{Workshop}: Repair or create simple items between adventures.
\item \textbf{Trading Post}: Acquire common goods at better prices.
\end{itemize}

\paragraph{Boon Activation}
Spend \textbf{1 Boon} to use an asset dramatically during a scene:
\begin{itemize}
\item \textbf{Safehouse}: Suddenly reveal a hidden escape route.
\item \textbf{Spy Network}: Produce crucial information at a critical moment.
\item \textbf{Workshop}: Create an improvised solution to an immediate problem.
\item \textbf{Trading Post}: Call in a favor from a business contact.
\end{itemize}

\paragraph{XP Activation}
Spend \textbf{2 XP} to use an asset's off-screen effect outside your normal allowance:
\begin{itemize}
\item Emergency use when you've already used your free activation.
\item Additional uses during downtime periods.
\item Special circumstances requiring extra asset support.
\end{itemize}

\section{Asset Conditions}
\index{assets!conditions}

Assets have condition states affecting their usefulness:

\subsection*{Condition Levels}
\begin{description}
\item[Maintained] Fully functional, no penalties.
\item[Neglected] $-1$ die when used; requires attention.
\item[Compromised] Unavailable until repaired or recovered.
\end{description}

\subsection*{Maintenance Requirements}
\begin{itemize}
\item \textbf{Regular Upkeep}: Pay XP equal to asset value each downtime period.
\item \textbf{Alternative}: Spend a scene tending to the asset.
\item \textbf{Neglect}: Assets deteriorate if not maintained.
\item \textbf{Recovery}: Compromised assets require significant effort to restore.
\end{itemize}

\section{Followers System}
\index{followers}

Followers are characters who assist you directly.

\subsection*{Follower Capability Ratings}
Followers are rated by Capability (\textbf{Cap}) from 1 to 5:
\begin{center}
\small
\begin{tabular}{ll}
\toprule
\textbf{Cap} & \textbf{Description} \\
\midrule
1 & Novice helper, basic assistance \\
2 & Competent assistant, reliable support \\
3 & Skilled specialist, valuable aid \\
4 & Expert ally, significant capability \\
5 & Master companion, exceptional ability \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Follower Costs}
\begin{itemize}
\item \textbf{XP Cost}: Capability squared (\emph{Cap}$^2$).
\item \textbf{Example}: Cap 3 follower costs $3^2=9$ XP.
\item \textbf{Recruitment}: 1--3 days downtime to find and brief.
\item \textbf{Limits}: The GM may set maximum followers based on story.
\end{itemize}

\subsection*{Follower Types}
\begin{description}
\item[Combat Allies] Warriors, guards, mercenaries.
\item[Technical Experts] Craftspeople, engineers, specialists.
\item[Social Contacts] Informants, diplomats, agents.
\item[Specialists] Unique capabilities like magic or stealth.
\end{description}

\section{Using Followers}
\index{followers!usage}

\subsection*{Assistance in Scenes}
Followers can help with your actions:
\begin{itemize}
\item \textbf{Assist Dice}: Add dice equal to $\min(\text{Cap}, \text{relevant skill})$.
\item \textbf{Maximum Bonus}: +3 dice total from all sources.
\item \textbf{Exception}: \emph{Exceptional Coordination} talent allows +4 from one follower.
\item \textbf{One Helper}: Only one follower can assist per action.
\end{itemize}

\subsection*{Independent Actions}
Once per scene (party-wide), a follower can take a small action:
\begin{description}
\item[\emph{Scout \& Signal}] Change an ally's next action to \textbf{Controlled} position.
\item[\emph{Distract \& Draw}] Reduce a threat clock by 1 segment.
\item[\emph{Fetch \& Carry}] Move an object through danger safely.
\end{description}

\subsection*{Cost of Independent Actions}
\begin{itemize}
\item Mark +1 \textbf{Exposure} (attention or stress), \emph{or}
\item Take \textbf{Harm 1} (injury or trauma).
\item Cannot be used if the follower is already \textbf{Compromised}.
\end{itemize}

\section{Follower Conditions}
\index{followers!conditions}

Followers track two condition types:

\subsection*{Exposure}
Represents attention, stress, or narrative pressure:
\begin{itemize}
\item \textbf{Gains}: From independent actions, dangerous situations, complications.
\item \textbf{Effects}: Increased risk, reduced effectiveness, attention from enemies.
\item \textbf{Recovery}: Downtime activities, careful management.
\end{itemize}

\subsection*{Harm}
Represents injury, trauma, or damage:
\begin{itemize}
\item \textbf{Gains}: From combat, accidents, enemy attacks.
\item \textbf{Effects}: Penalties to assistance, possible incapacity.
\item \textbf{Recovery}: Medical care, rest, magical healing.
\end{itemize}

\subsection*{Condition States}
\begin{description}
\item[Maintained] Ready and reliable, full capability.
\item[Neglected] Needs attention, $-1$ die to assistance.
\item[Compromised] Unavailable: captured, defected, lost, or incapacitated.
\end{description}

\section{Follower Risks}
\index{followers!risks}

Using followers carries significant risks:

\subsection*{Complication Targeting}
When the GM spends \textbf{2+ Story Beats} on an action where you have assistance:
\begin{itemize}
\item The follower may face consequences instead of you.
\item Could be injury, capture, betrayal, or other complications.
\item Fictionally appropriate to the situation.
\end{itemize}

\subsection*{Off-Screen Capability}
Once per downtime, a \textbf{Cap 5} follower can solve a significant problem:
\begin{itemize}
\item But generates \textbf{1 Story Beat} for the party.
\item The GM describes how this creates story consequences.
\item Useful for emergencies but costly.
\end{itemize}

\section{Upkeep and Maintenance}
\index{upkeep}

Both assets and followers require regular maintenance.

\subsection*{Asset Upkeep}
\begin{itemize}
\item \textbf{XP Cost}: Equal to asset value each downtime.
\item \textbf{Scene Alternative}: Spend a scene managing the asset.
\item \textbf{Neglect}: Assets deteriorate to \emph{Neglected}, then \emph{Compromised}.
\item \textbf{Recovery}: Compromised assets require significant investment.
\end{itemize}

\subsection*{Follower Upkeep}
\begin{itemize}
\item \textbf{XP Cost}: Equal to follower's \emph{Cap} each downtime.
\item \textbf{Scene Alternative}: Spend a scene maintaining the relationship.
\item \textbf{Loyalty}: Followers may leave if neglected or mistreated.
\item \textbf{Recovery}: Compromised followers require rescue or reconciliation.
\end{itemize}

\section{Strategic Considerations}
\index{strategic considerations}

\subsection*{When to Invest in Assets}
\begin{itemize}
\item You need reliable off-screen capabilities.
\item Your character concept involves wealth or influence.
\item The party lacks certain logistical support.
\item You want to build long-term influence.
\end{itemize}

\subsection*{When to Invest in Followers}
\begin{itemize}
\item You need on-screen assistance.
\item Your character works better with support.
\item The party needs specific capabilities you lack.
\item You want character-driven story opportunities.
\end{itemize}

\subsection*{Balance Recommendations}
\begin{itemize}
\item \textbf{Personal Path}: 0--10\% assets/followers.
\item \textbf{Balanced Path}: 15--25\% assets/followers.
\item \textbf{Influencer Path}: 35--55\% assets/followers.
\end{itemize}

\section{Loyalty and Relationships}
\index{loyalty}\index{relationships}

\subsection*{Loyalty Levels}
Optional system for tracking follower loyalty:
\begin{description}
\item[Wary] Cautious, may leave if pressured; +1 XP upkeep cost.
\item[Steady] Reliable, standard performance; normal upkeep.
\item[Devoted] Loyal, may sacrifice; can convert one major complication to a minor setback per arc.
\end{description}

\subsection*{Building Loyalty}
\begin{itemize}
\item Fair treatment and respect.
\item Sharing rewards and successes.
\item Protecting followers from harm.
\item Honoring agreements and promises.
\end{itemize}

\subsection*{Losing Loyalty}
\begin{itemize}
\item Mistreatment or disrespect.
\item Unreasonable demands or risks.
\item Broken promises or betrayal.
\item Consistent neglect.
\end{itemize}

\section{Advanced Follower Management}
\index{followers!advanced}

\subsection*{Follower Groups}
For multiple similar followers, you can manage them as a group:
\begin{itemize}
\item \textbf{Single Rating}: Treat as one entity with combined capability.
\item \textbf{Condition Tracking}: Group shares exposure and harm.
\item \textbf{Maintenance}: Single upkeep cost for the group.
\item \textbf{Risks}: Problems affect the entire group.
\end{itemize}

\subsection*{Follower Advancement}
Followers can improve over time:
\begin{itemize}
\item \textbf{Experience}: Gain capability through successful assistance.
\item \textbf{Training}: Spend XP to improve follower capabilities.
\item \textbf{Equipment}: Better gear can enhance effectiveness.
\item \textbf{Limits}: Followers typically cap at lower levels than PCs.
\end{itemize}

\section{Risk Management}
\index{risk management}

\subsection*{Asset Risks}
\begin{itemize}
\item \textbf{Financial}: Assets can be costly to maintain.
\item \textbf{Security}: Assets can be attacked or stolen.
\item \textbf{Attention}: Valuable assets draw unwanted notice.
\item \textbf{Dependency}: Over-reliance can be problematic.
\end{itemize}

\subsection*{Follower Risks}
\begin{itemize}
\item \textbf{Safety}: Followers can be harmed or captured.
\item \textbf{Loyalty}: Followers may betray or leave.
\item \textbf{Attention}: Followers can draw enemy interest.
\item \textbf{Morale}: Followers have needs and limits.
\end{itemize}

\subsection*{Mitigation Strategies}
\begin{itemize}
\item \textbf{Diversification}: Don't put all resources in one place.
\item \textbf{Security}: Protect valuable assets and followers.
\item \textbf{Relationships}: Maintain good terms with your people.
\item \textbf{Contingencies}: Have backup plans for losses.
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Assets and Followers Quick Reference,fonttitle=\bfseries]
\textbf{Assets:}
\begin{itemize}
\item Minor: 4 XP \;|\; Standard: 8 XP \;|\; Major: 12 XP
\item Free off-screen use: once per session
\item Boon activation: spend 1 Boon for scene impact
\item Conditions: \emph{Maintained} $\rightarrow$ \emph{Neglected} $\rightarrow$ \emph{Compromised}
\end{itemize}

\textbf{Followers:}
\begin{itemize}
\item Cost: \emph{Cap}$^2$ XP
\item Assistance: $+\min(\text{Cap},\text{skill})$ dice (max +3 from all sources)
\item Independent action: once per scene (party-wide)
\item Conditions: \emph{Exposure} and \emph{Harm} tracks
\end{itemize}

\textbf{Upkeep:} Pay XP or spend a scene each downtime
\end{tcolorbox}

\section{Practical Examples}

\subsection*{Asset Example: The Safehouse}
\begin{itemize}
\item \textbf{Type}: Minor Asset (4 XP)
\item \textbf{Free Use}: Secure lodging, basic supplies between adventures.
\item \textbf{Boon Activation}: Reveal a hidden escape route during pursuit.
\item \textbf{Upkeep}: 4 XP or one scene each downtime.
\item \textbf{Risks}: Discovery by enemies, maintenance costs.
\end{itemize}

\subsection*{Follower Example: The Scout}
\begin{itemize}
\item \textbf{Capability}: 3 (9 XP cost)
\item \textbf{Assistance}: +3 dice on tracking and survival rolls.
\item \textbf{Independent Action}: Scout ahead to improve party position.
\item \textbf{Upkeep}: 3 XP or relationship scene each downtime.
\item \textbf{Risks}: Injury in dangerous scouting; disloyalty if mistreated.
\end{itemize}

\subsection*{Combination Example: The Merchant}
\begin{itemize}
\item \textbf{Assets}: Trading post (8 XP), caravan (4 XP) --- \emph{12 XP total}
\item \textbf{Followers}: Cap 2 guards \textbf{(two)} (4 XP each = 8 XP), Cap 3 factor (9 XP) --- \emph{17 XP total}
\item \textbf{Total Investment}: \textbf{29 XP} in assets and followers
\item \textbf{Upkeep}: Assets 12 XP + Followers (2+2+3 = 7 XP) = \textbf{19 XP} per downtime period
\item \textbf{Benefits}: Trade income, transport, protection, business contacts
\item \textbf{Risks}: Competition, bandit attacks, employee issues, regulatory attention
\end{itemize}

\paragraph{Note on Upkeep Math:}
If you instead hire \emph{one} Cap 2 guard (4 XP) and keep the Cap 3 factor (9 XP), follower upkeep becomes $2+3=5$ and total upkeep becomes $12+5=\textbf{17}$ per downtime.

Remember: Assets and followers can greatly expand your capabilities, but they require careful management and carry significant risks. Invest wisely based on your character concept and the needs of your group.

\section{Narrative-Heavy Asset and Follower Options}

For groups that prefer strong narrative focus in asset and follower management, consider these optional approaches:

\textbf{Story-Driven Upkeep:} Instead of tracking XP costs for upkeep, the GM can introduce narrative complications that require attention. A neglected asset might attract unwanted attention, while a neglected follower might request a favor or special treatment.

\textbf{Collaborative Management:} Players can describe how they maintain their assets and followers through roleplay rather than mechanical upkeep costs. A well-described scene of tending to a workshop or bonding with followers can fulfill maintenance requirements.

\textbf{Asset and Follower as Character Development:} Use asset and follower management as opportunities for character growth and backstory development, allowing players to narrate how their relationships and holdings evolve through significant story moments.

\textbf{Flexible Condition Tracking:} Focus on the narrative implications of asset and follower conditions rather than strict mechanical penalties. A "Neglected" asset might still function but with interesting complications, while a "Compromised" asset might require creative solutions rather than just XP investment.\end{chapter}