\chapter{Gods, Powers, and Patrons}
\label{ch:gods-powers-patrons}

\index{gods}\index{powers}\index{patrons}\index{rites}\index{obligation}
In \textit{Fate's Edge}, the \textbf{divine} is not a distant thesis—it is an active pressure system that pushes back when named. Temples argue in \emph{Ecktic High} (\S\ref{ch:languages}), bells speak law in the Mistlands, and oaths sworn under clear sky bind tighter than iron (\S\ref{ch:world-regions}). This chapter restores the \textbf{lore of traditions}, clarifies \textbf{Patrons} and their \textbf{Rites}, and shows how all of it interlocks with the core systems (\S\ref{ch:magic}).

\section{The Nature of Divine Forces}
\index{divine forces}\index{numina}\index{offices (divine)}
Scholars of Thepyrgos write that divinity condenses in three ways:
\begin{description}
  \item[\textbf{Numina}] \index{numina} Place-bound or concept-focused powers (a harbor's luck, a winter's severity).
  \item[\textbf{Lineages}] \index{lineages!divine} Ancestor-chains and civic cults that accrue obligation over centuries.
  \item[\textbf{Offices}] \index{offices (divine)!seats} Roles that exist whether or not a face is seated—\emph{Witness}, \emph{Arbiter}, \emph{Wayfinder}. Whoever answers the call \emph{is} the office for as long as the rites hold.
\end{description}
All three \emph{answer} when approached in their proper language, place, and price (\S\ref{ch:languages}).

\paragraph{Friction With Reality.}\index{backlash}\index{Story Beats (SB)}
Calling on the sacred generates narrative pressure. Treat divine missteps as Story Beats (SB) that the GM can spend for omens, visitations, taboos invoked, or social consequences among the faithful (\S\ref{ch:magic}).

\section{Major Divine Traditions}
\index{divine traditions}\index{cults}\index{temples}

\subsection*{The Everflame and the Lampers}
\index{Everflame}\index{Lampers@Lampers (urban confraternities)}
\textbf{Domains:} Fire, purification, law \quad
\textbf{Centers:} Ecktoria's marble courts, Sun Coast porticoes \\
\textbf{Themes:} Order through light, confession by heat, civic duty
\begin{itemize}
  \item \textbf{Clergy \& Orders:} \emph{Lampers} tend beacon-fires and court lamps; \emph{Ash-Deacons} oversee penances. \index{orders!Lampers}\index{orders!Ash-Deacons}
  \item \textbf{Rites:} \emph{Lamp of Plain Speech} (forces truth in sworn testimony), \emph{Cinder-Ward} (burns contagion, literal or social).
  \item \textbf{Customs:} \emph{Three-Wick Vigil} before verdicts; extinguishing a lamp early is a public insult. \index{customs!three-wick vigil}
\end{itemize}
\emph{Vignette:} \textit{In the Echo Hall a thousand glass panes bloom with flame. Each verdict adds one. The Archivist snuffs two and the crowd inhales—history itself has been corrected.}

\subsection*{The Reformed Light (Mercy-Cant)}
\index{Light Reformation}\index{mercy-cant}
\textbf{Domains:} Revelation, mercy, community \quad
\textbf{Centers:} Valewood clinics, Theonan schools \\
\textbf{Themes:} Repair over punishment; unity through service
\begin{itemize}
  \item \textbf{Orders:} \emph{Lantern-Hands} (healers), \emph{Keepers of Account} (community ledgers of generosity). \index{orders!Lantern-Hands}\index{orders!Keepers of Account}
  \item \textbf{Rites:} \emph{Sharing-Bread} (multiplies supplies within a bounded fellowship), \emph{Pardon-Mark} (removes a social stigma clock if restitution is made).
  \item \textbf{Taboo:} No light is to be used to shame the poor in public.
\end{itemize}

\subsection*{Sea and Storm Courts}
\index{Sea and Storm}\index{courts!maritime}
\textbf{Domains:} Oceans, storms, navigation \quad
\textbf{Centers:} Linn and Zakov archipelagos (\S\ref{ch:world-regions}) \\
\textbf{Themes:} Bargain with weather; respect for hazard
\begin{itemize}
  \item \textbf{Priesthood:} \emph{Tide-Readers} interpret current-scripts; \emph{Bell-Helms} lead weather-rites. \index{orders!Tide-Readers}\index{orders!Bell-Helms}
  \item \textbf{Rites:} \emph{Stern-Knot} (locks a course against drift), \emph{Storm's Portion} (you surrender cargo; the squall spares lives).
  \item \textbf{Language:} Sea-Patter requires triadic call-and-response (\S\ref{ch:languages}); missing the third call risks an omen. \index{Sea-Patter}
\end{itemize}

\subsection*{Sky and Conflict Traditions}
\index{Sky}\index{Conflict}
\textbf{Domains:} Sky, conflict, honor \quad
\textbf{Centers:} Vilikari highlands, Acasian Marches \\
\textbf{Themes:} Oaths make war bearable; victory is \emph{clean} when witnessed
\begin{itemize}
  \item \textbf{Oath-Forms:} Horn-codes and duel formulae (\S\ref{ch:languages}) bind reparations. \index{horn-codes}
  \item \textbf{Rites:} \emph{Sky-Witness} (freezes a parley one exchange; no one may strike), \emph{Banner-Balance} (equalizes a skirmish's scale for one beat).
\end{itemize}

\subsection*{Shadow and Transformation Paths}
\index{Shadow}\index{Transformation}
\textbf{Domains:} Shade, secrets, change \quad
\textbf{Centers:} Eastern caravan cloisters, Mistland bell precincts \\
\textbf{Themes:} Hidden knowledge; necessary metamorphosis
\begin{itemize}
  \item \textbf{Initiates:} \emph{Veil-Bearers} trade names seasonally (see Fog-Names, \S\ref{ch:languages}). \index{Veil-Bearers}\index{Fog-Names}
  \item \textbf{Rites:} \emph{Veil of Change} (mask a social role until challenged by bell or oath), \emph{Moth-Path} (take a consequence now to step past a danger later).
\end{itemize}

\subsection*{Ancestral Houses and Stone-Law}
\index{Ancestral veneration}\index{Stone-Law}
\textbf{Domains:} Kin, craft, memory \quad
\textbf{Centers:} Stone Kingdoms, Aeler tunnels \\
\textbf{Themes:} Work remembered; promises become architecture
\begin{itemize}
  \item \textbf{Practices:} Charter-Script read aloud to attest memory (\S\ref{ch:languages}). \index{Charter-Script}
  \item \textbf{Rites:} \emph{Line-Recall} (perfect recall in spaces your kin laid), \emph{Load-Share} (shift harm from structure to sworn caretakers as Fatigue).
\end{itemize}

\section{Patrons and the Rites}
\index{patrons}\index{rites}\index{obligation}\index{attention}
Not all power wants worship. \textbf{Patrons} are entities or \emph{offices} that answer to named procedures and ledgered \textbf{Obligation}. Some wear faces. Others are the role itself. Runekeepers pledge themselves to a single Patron, Invokers use Symbols and arcane knowledge to perform their rites. 

\subsection*{Devotional Practice}
\index{devotion!practice}
\emph{Prayer, calendar rites, ethics.} Grants access to themed \textbf{Talents} (e.g., \emph{Purifying Touch}, \emph{Guide's Wisdom}) at normal XP cost (\S\ref{ch:talents}). Use language, posture, and season for fictional advantage (\S\ref{ch:languages}).

\subsection*{Thaumaturgical Access}
\index{thaumaturgy}
Treat miracles as procedures: preconditions, components, and test points. Use \emph{Wits + Arcana} to \emph{Channel}, then \emph{Weave} (\S\ref{ch:magic}); divine sources may swap in \emph{Spirit + Faith} for either phase, per tradition.

\subsection*{Pact Magic (Rites)}
\index{pact magic}\index{rites!invocation}
\begin{itemize}
  \item \textbf{Prereqs:} Talent such as \emph{Familiar Bond} or \emph{Codex of Names}. \index{talents!Familiar Bond}\index{talents!Codex of Names}
  \item \textbf{Invoke:} Spend a \emph{Boon} \emph{or} mark \emph{Obligation} (GM sets segments based on scale: minor 1, standard 2, major 3+).
  \item \textbf{Risk:} On misses, GM spends SB as \emph{Backlash} flavored by the Patron's office (\S\ref{ch:magic}).
\end{itemize}

\section{Regional Faces \& Sacred Geography}
\index{regional religion}\index{sacred sites}\index{pilgrimage}

\subsection*{Inland Sea Circuit}
\index{Inland Sea}
\textbf{Pilgrim's Ring}—a route of beacon-towers, courts, and granaries. Completing the ring in a year and a day grants the title \emph{Road-Kin} (merchants will open a door once for free). \index{Road-Kin}

\subsection*{Mistlands Bell-Precincts}
\index{Mistlands!bells}\index{Bell-Speech}
Bells carry law and lament. The \emph{Three-Tone Dusk} is binding: certain Rites (\emph{Veil of Change}) must answer or disperse harmlessly.

\subsection*{Stone Kingdoms Oath-Works}
\index{Stone Kingdoms}\index{Stone-Speech}
Bridges and vaults are "signed" with ancestor marks. A \emph{Line-Recall} Rite in the right hall reads generations like a book.

\subsection*{Kahfagian Sea-Courts}
\index{Kahfagian}\index{Sea courts}
Courts convene on the sea itself: three boats, three judges, verdict by current and chant (\S\ref{ch:languages} Sea-Patter).

\section{Talents \& Devotional Boons}
\index{talents!divine}\index{boons!divine}
Examples that fit the above traditions (costs per \S\ref{ch:talents}):
\begin{itemize}
  \item \textbf{Purifying Touch}—Once per scene, cleanse a mundane taint or calm a crowd segment if you carry fire openly.
  \item \textbf{Guide's Wisdom}—When mediating, ask the GM one clarifying question about \emph{stakes}; answer is candid.
  \item \textbf{Storm Sense}—You always know the fastest safe anchorage within a day's sail.
  \item \textbf{Oath-Keeper}—When you publicly accept terms, gain a \emph{Boon} to the next roll that advances them.
\end{itemize}

\section{Consequences \& Conflicts}
\index{consequences!divine}\index{conflict!religious}
\begin{itemize}
  \item \textbf{Attention:} High Obligation grants visitations—aid that arrives \emph{with} a demand.
  \item \textbf{Rival Doctrines:} Lampers vs Mercy-Cant on punishment; Sea Courts vs Sky Oath on jurisdiction; Ancestral Houses vs Catalyst on continuity.
  \item \textbf{Blasphemy as Mechanics:} GM spends SB to flip a rite's tone (e.g., \emph{Lamp of Plain Speech} exposes \emph{your} omission first).
\end{itemize}

\section{Gameplay Tools}
\index{GM tools!divine}\index{oracles!religious}
\subsection*{Quick SB Menu (Divine)}
\begin{itemize}
  \item \textbf{1 SB}: An omen complicates timing or tone (bell toll, lamp gutter).
  \item \textbf{2 SB}: A taboo is invoked; someone important is offended.
  \item \textbf{3 SB}: A rival office answers as well (Keeper \emph{vs} Catalyst).
  \item \textbf{4 SB}: An owed service comes due \emph{now}.
\end{itemize}

\subsection*{Sacred Site Tags}
\index{sacred sites!tags}
\textit{Consecrated}, \textit{Witnessed}, \textit{Thresholded}, \textit{Storm-Favored}, \textit{Ancestral-Resonant}. Each tag lowers DV by 1 for aligned rites and raises it by 1 for opposed offices.

\begin{tcolorbox}[colback=purple!5!white,colframe=purple!75!black,title=Divine \& Patron Quick Reference,fonttitle=\bfseries]
\textbf{Approaches:} Devotion (Talents), Thaumaturgy (procedures), Pact (Rites + Obligation) \\
\textbf{Obligation Track:} 9 segments—at 3/6/9 trigger requests, visitations, demands \\
\textbf{Backlash:} GM SB spend, flavored by office and region \\
\textbf{Language Keys:} Ecktic High (oaths), Sea-Patter (weather), Charter-Script (memory) \\
\textbf{Assets Help:} Libraries, colleges, shrines modify DV/Position (\S\ref{ch:compendium-assets})
\end{tcolorbox}

\section{Mechanical Integration}
\index{mechanics!divine}\index{rites!mechanics}
\begin{itemize}
  \item \textbf{Talent Access}—Divine-themed Talents use normal XP (\S\ref{ch:talents}); prerequisites are narrative.
  \item \textbf{Rites System}—Use \emph{Obligation} and SB as described; scaling: minor (1 seg), standard (2), major (3+).
  \item \textbf{Casting Loop}—Rites can slot into \emph{Channel/Weave/Backlash} (\S\ref{ch:magic}); \emph{Spirit} often substitutes for \emph{Wits}.
  \item \textbf{Language Hooks}—Correct register can lower DV by 1 or grant \emph{Boon} (\S\ref{ch:languages}).
\end{itemize}

\section*{Closing Note}
\index{allegiance}\index{story hooks}
Every allegiance writes \emph{terms} into your story: who will answer when you call, and who will come to collect when you don't.