\chapter{Gods, Powers, and Patrons}
\label{ch:gods-powers-patrons}

\index{gods}\index{powers}\index{patrons}\index{rites}\index{obligation}
In \textit{Fate's Edge}, the \textbf{divine} is not a distant thesis—it is an active pressure system that pushes back when named. Temples argue in \emph{Ecktic High} (\S\ref{ch:languages}), bells speak law in the Mistlands, and oaths sworn under clear sky bind tighter than iron (\S\ref{ch:world-regions}). This chapter restores the \textbf{lore of traditions}, clarifies \textbf{Patrons} and their \textbf{Rites}, and shows how all of it interlocks with the core systems (\S\ref{ch:magic}).

\section{The Nature of Divine Forces}
\index{divine forces}\index{numina}\index{offices (divine)}
Scholars of Thepyrgos write that divinity condenses in three ways:
\begin{description}
  \item[\textbf{Numina}] \index{numina} Place-bound or concept-focused powers (a harbor's luck, a winter's severity).
  \item[\textbf{Lineages}] \index{lineages!divine} Ancestor-chains and civic cults that accrue obligation over centuries.
  \item[\textbf{Offices}] \index{offices (divine)!seats} Roles that exist whether or not a face is seated—\emph{Witness}, \emph{Arbiter}, \emph{Wayfinder}. Whoever answers the call \emph{is} the office for as long as the rites hold.
\end{description}
All three \emph{answer} when approached in their proper language, place, and price (\S\ref{ch:languages}).

\paragraph{Friction With Reality.}\index{backlash}\index{Story Beats (SB)}
Calling on the sacred generates narrative pressure. Treat divine missteps as Story Beats (SB) that the GM can spend for omens, visitations, taboos invoked, or social consequences among the faithful (\S\ref{ch:magic}).

\section{Major Divine Traditions}
\index{divine traditions}\index{cults}\index{temples}

\subsection*{The Everflame and the Lampers}
\index{Everflame}\index{Lampers@Lampers (urban confraternities)}
\textbf{Domains:} Fire, purification, law \quad
\textbf{Centers:} Ecktoria's marble courts, Sun Coast porticoes \\
\textbf{Themes:} Order through light, confession by heat, civic duty
\begin{itemize}
  \item \textbf{Clergy \& Orders:} \emph{Lampers} tend beacon-fires and court lamps; \emph{Ash-Deacons} oversee penances. \index{orders!Lampers}\index{orders!Ash-Deacons}
  \item \textbf{Rites:} \emph{Lamp of Plain Speech} (forces truth in sworn testimony), \emph{Cinder-Ward} (burns contagion, literal or social).
  \item \textbf{Customs:} \emph{Three-Wick Vigil} before verdicts; extinguishing a lamp early is a public insult. \index{customs!three-wick vigil}
\end{itemize}
\emph{Vignette:} \textit{In the Echo Hall a thousand glass panes bloom with flame. Each verdict adds one. The Archivist snuffs two and the crowd inhales—history itself has been corrected.}

\subsection*{The Reformed Light (Mercy-Cant)}
\index{Light Reformation}\index{mercy-cant}
\textbf{Domains:} Revelation, mercy, community \quad
\textbf{Centers:} Valewood clinics, Theonan schools \\
\textbf{Themes:} Repair over punishment; unity through service
\begin{itemize}
  \item \textbf{Orders:} \emph{Lantern-Hands} (healers), \emph{Keepers of Account} (community ledgers of generosity). \index{orders!Lantern-Hands}\index{orders!Keepers of Account}
  \item \textbf{Rites:} \emph{Sharing-Bread} (multiplies supplies within a bounded fellowship), \emph{Pardon-Mark} (removes a social stigma clock if restitution is made).
  \item \textbf{Taboo:} No light is to be used to shame the poor in public.
\end{itemize}

\subsection*{Sea and Storm Courts}
\index{Sea and Storm}\index{courts!maritime}
\textbf{Domains:} Oceans, storms, navigation \quad
\textbf{Centers:} Linn and Zakov archipelagos (\S\ref{ch:world-regions}) \\
\textbf{Themes:} Bargain with weather; respect for hazard
\begin{itemize}
  \item \textbf{Priesthood:} \emph{Tide-Readers} interpret current-scripts; \emph{Bell-Helms} lead weather-rites. \index{orders!Tide-Readers}\index{orders!Bell-Helms}
  \item \textbf{Rites:} \emph{Stern-Knot} (locks a course against drift), \emph{Storm's Portion} (you surrender cargo; the squall spares lives).
  \item \textbf{Language:} Sea-Patter requires triadic call-and-response (\S\ref{ch:languages}); missing the third call risks an omen. \index{Sea-Patter}
\end{itemize}

\subsection*{Sky and Conflict Traditions}
\index{Sky}\index{Conflict}
\textbf{Domains:} Sky, conflict, honor \quad
\textbf{Centers:} Vilikari highlands, Acasian Marches \\
\textbf{Themes:} Oaths make war bearable; victory is \emph{clean} when witnessed
\begin{itemize}
  \item \textbf{Oath-Forms:} Horn-codes and duel formulae (\S\ref{ch:languages}) bind reparations. \index{horn-codes}
  \item \textbf{Rites:} \emph{Sky-Witness} (freezes a parley one exchange; no one may strike), \emph{Banner-Balance} (equalizes a skirmish's scale for one beat).
\end{itemize}

\subsection*{Shadow and Transformation Paths}
\index{Shadow}\index{Transformation}
\textbf{Domains:} Shade, secrets, change \quad
\textbf{Centers:} Eastern caravan cloisters, Mistland bell precincts \\
\textbf{Themes:} Hidden knowledge; necessary metamorphosis
\begin{itemize}
  \item \textbf{Initiates:} \emph{Veil-Bearers} trade names seasonally (see Fog-Names, \S\ref{ch:languages}). \index{Veil-Bearers}\index{Fog-Names}
  \item \textbf{Rites:} \emph{Veil of Change} (mask a social role until challenged by bell or oath), \emph{Moth-Path} (take a consequence now to step past a danger later).
\end{itemize}

\subsection*{Ancestral Houses and Stone-Law}
\index{Ancestral veneration}\index{Stone-Law}
\textbf{Domains:} Kin, craft, memory \quad
\textbf{Centers:} Stone Kingdoms, Aeler tunnels \\
\textbf{Themes:} Work remembered; promises become architecture
\begin{itemize}
  \item \textbf{Practices:} Charter-Script read aloud to attest memory (\S\ref{ch:languages}). \index{Charter-Script}
  \item \textbf{Rites:} \emph{Line-Recall} (perfect recall in spaces your kin laid), \emph{Load-Share} (shift harm from structure to sworn caretakers as Fatigue).
\end{itemize}

\section{Patrons and the Rites}
\index{patrons}\index{rites}\index{obligation}\index{attention}
Not all power wants worship. \textbf{Patrons} are entities or \emph{offices} that answer to named procedures and ledgered \textbf{Obligation}. Some wear faces. Others are the role itself. Runekeepers pledge themselves to a single Patron, Invokers use Symbols and arcane knowledge to perform their rites. 

\subsection*{Devotional Practice}
\index{devotion!practice}
\emph{Prayer, calendar rites, ethics.} Grants access to themed \textbf{Talents} (e.g., \emph{Purifying Touch}, \emph{Guide's Wisdom}) at normal XP cost (\S\ref{ch:talents}). Use language, posture, and season for fictional advantage (\S\ref{ch:languages}).

\subsection*{Thaumaturgical Access}
\index{thaumaturgy}
Treat miracles as procedures: preconditions, components, and test points. Use \emph{Wits + Arcana} to \emph{Channel}, then \emph{Weave} (\S\ref{ch:magic}); divine sources may swap in \emph{Spirit + Faith} for either phase, per tradition.

\subsection*{Pact Magic (Rites)}
\index{pact magic}\index{rites!invocation}
\begin{itemize}
  \item \textbf{Prereqs:} Talent such as \emph{Familiar Bond} or \emph{Codex of Names}. \index{talents!Familiar Bond}\index{talents!Codex of Names}
  \item \textbf{Invoke:} Spend a \emph{Boon} \emph{or} mark \emph{Obligation} (GM sets segments based on scale: minor 1, standard 2, major 3+).
  \item \textbf{Risk:} On misses, GM spends SB as \emph{Backlash} flavored by the Patron's office (\S\ref{ch:magic}).
\end{itemize}

\section{Regional Faces \& Sacred Geography}
\index{regional religion}\index{sacred sites}\index{pilgrimage}

\subsection*{Inland Sea Circuit}
\index{Inland Sea}
\textbf{Pilgrim's Ring}—a route of beacon-towers, courts, and granaries. Completing the ring in a year and a day grants the title \emph{Road-Kin} (merchants will open a door once for free). \index{Road-Kin}

\subsection*{Mistlands Bell-Precincts}
\index{Mistlands!bells}\index{Bell-Speech}
Bells carry law and lament. The \emph{Three-Tone Dusk} is binding: certain Rites (\emph{Veil of Change}) must answer or disperse harmlessly.

\subsection*{Stone Kingdoms Oath-Works}
\index{Stone Kingdoms}\index{Stone-Speech}
Bridges and vaults are "signed" with ancestor marks. A \emph{Line-Recall} Rite in the right hall reads generations like a book.

\subsection*{Kahfagian Sea-Courts}
\index{Kahfagian}\index{Sea courts}
Courts convene on the sea itself: three boats, three judges, verdict by current and chant (\S\ref{ch:languages} Sea-Patter).

\section{Talents \& Devotional Boons}
\index{talents!divine}\index{boons!divine}
Examples that fit the above traditions (costs per \S\ref{ch:talents}):
\begin{itemize}
  \item \textbf{Purifying Touch}—Once per scene, cleanse a mundane taint or calm a crowd segment if you carry fire openly.
  \item \textbf{Guide's Wisdom}—When mediating, ask the GM one clarifying question about \emph{stakes}; answer is candid.
  \item \textbf{Storm Sense}—You always know the fastest safe anchorage within a day's sail.
  \item \textbf{Oath-Keeper}—When you publicly accept terms, gain a \emph{Boon} to the next roll that advances them.
\end{itemize}

\section{Consequences \& Conflicts}
\index{consequences!divine}\index{conflict!religious}
\begin{itemize}
  \item \textbf{Attention:} High Obligation grants visitations—aid that arrives \emph{with} a demand.
  \item \textbf{Rival Doctrines:} Lampers vs Mercy-Cant on punishment; Sea Courts vs Sky Oath on jurisdiction; Ancestral Houses vs Catalyst on continuity.
  \item \textbf{Blasphemy as Mechanics:} GM spends SB to flip a rite's tone (e.g., \emph{Lamp of Plain Speech} exposes \emph{your} omission first).
\end{itemize}

\section{Gameplay Tools}
\index{GM tools!divine}\index{oracles!religious}
\subsection*{Quick SB Menu (Divine)}
\begin{itemize}
  \item \textbf{1 SB}: An omen complicates timing or tone (bell toll, lamp gutter).
  \item \textbf{2 SB}: A taboo is invoked; someone important is offended.
  \item \textbf{3 SB}: A rival office answers as well (Keeper \emph{vs} Catalyst).
  \item \textbf{4 SB}: An owed service comes due \emph{now}.
\end{itemize}

\subsection*{Sacred Site Tags}
\index{sacred sites!tags}
\textit{Consecrated}, \textit{Witnessed}, \textit{Thresholded}, \textit{Storm-Favored}, \textit{Ancestral-Resonant}. Each tag lowers DV by 1 for aligned rites and raises it by 1 for opposed offices.

\begin{tcolorbox}[colback=purple!5!white,colframe=purple!75!black,title=Divine \& Patron Quick Reference,fonttitle=\bfseries]
\textbf{Approaches:} Devotion (Talents), Thaumaturgy (procedures), Pact (Rites + Obligation) \\
\textbf{Obligation Track:} 9 segments—at 3/6/9 trigger requests, visitations, demands \\
\textbf{Backlash:} GM SB spend, flavored by office and region \\
\textbf{Language Keys:} Ecktic High (oaths), Sea-Patter (weather), Charter-Script (memory) \\
\textbf{Assets Help:} Libraries, colleges, shrines modify DV/Position (\S\ref{ch:compendium-assets})
\end{tcolorbox}

\section{Mechanical Integration}
\index{mechanics!divine}\index{rites!mechanics}
\begin{itemize}
  \item \textbf{Talent Access}—Divine-themed Talents use normal XP (\S\ref{ch:talents}); prerequisites are narrative.
  \item \textbf{Rites System}—Use \emph{Obligation} and SB as described; scaling: minor (1 seg), standard (2), major (3+).
  \item \textbf{Casting Loop}—Rites can slot into \emph{Channel/Weave/Backlash} (\S\ref{ch:magic}); \emph{Spirit} often substitutes for \emph{Wits}.
  \item \textbf{Language Hooks}—Correct register can lower DV by 1 or grant \emph{Boon} (\S\ref{ch:languages}).
\end{itemize}

\section*{Closing Note}
\index{allegiance}\index{story hooks}
Every allegiance writes \emph{terms} into your story: who will answer when you call, and who will come to collect when you don't.

\clearpage
\section{Of Patrons, Runes, and Invokers}
\label{sec:invoker-lore}

\begin{quote}
``You wish to walk the road of power? Then listen well. The world is old, and older still are the voices beneath it. We call them \textit{Patrons}, though they were never sworn to us. They are the tides that move unseen, the keepers of forgotten bargains, the sleepers beneath the stone and the stars. To call upon them is to dip a hand into a river that has carved mountains.''
\end{quote}

\subsubsection{The Patrons}
\index{Patrons}

Patrons are vast intelligences---not gods, though some worship them as such. They are embodiments of \textit{concepts} and \textit{forces} rather than sovereigns. Raéyn, mistress of the tides and the sea-routes. Khemesh, the crushing inevitability of the deep. Nidhoggr, the worm that dreams in the roots of time. Each offers power, but always with cost: fatigue, scars upon fate, or a slow unweaving of one’s own story.

To entreat a Patron is to risk being marked. Their Rites are gifts and snares both.

% Individual patrons (order however you like)

\input{../reference/patrons/aveh.tex}
\clearpage
\input{../reference/patrons/carrion-king.tex}
\clearpage
\input{../reference/patrons/clockwork-monad.tex}
\clearpage
\input{../reference/patrons/confessor.tex}
\clearpage
\input{../reference/patrons/gallows-bell.tex}
\clearpage
\input{../reference/patrons/grimmir.tex}
\clearpage
\input{../reference/patrons/ikasha.tex}
\clearpage
\input{../reference/patrons/inaea.tex}
\clearpage
\input{../reference/patrons/inquisitor-prime.tex}
\clearpage
\input{../reference/patrons/isoka.tex}
\clearpage
\input{../reference/patrons/khemesh.tex}
\clearpage
\input{../reference/patrons/livaea.tex}
\clearpage
\input{../reference/patrons/lunara.tex}
\clearpage
\input{../reference/patrons/mab.tex}
\clearpage
\input{../reference/patrons/maelstraeus.tex}
\clearpage
\input{../reference/patrons/Malachai.tex}
\clearpage
\input{../reference/patrons/mir-iraath.tex}
\clearpage
\input{../reference/patrons/morag.tex}
\clearpage
\input{../reference/patrons/mykkiel.tex}
\clearpage
\input{../reference/patrons/nidhoggr.tex}
\clearpage
\input{../reference/patrons/ninth-rim.tex}
\clearpage
\input{../reference/patrons/oath-flame-light.tex}
\clearpage
\input{../reference/patrons/pale-shepherd.tex}
\clearpage
\input{../reference/patrons/palinode.tex}
\clearpage
\input{../reference/patrons/raeyn.tex}
\clearpage
\input{../reference/patrons/sacred-geometry.tex}
\clearpage
\input{../reference/patrons/sealed-gate.tex}
\clearpage
\input{../reference/patrons/silent-choir.tex}
\clearpage
\input{../reference/patrons/thrysos.tex}
\clearpage
\input{../reference/patrons/traveler.tex}
\clearpage
\input{../reference/patrons/varnek-karn.tex}
\clearpage
\input{patrons/vorthak.tex}
\clearpage
\input{../reference/patrons/witness.tex}
\clearpage
\input{../reference/patrons/xhak-thul.tex}
\clearpage

\paragraph{No True Acolyte}
Interpreting a patron's will is often a dangerous prospect in and of itself. Many a Runkeeper has found themselves on the opposite end of machinations from others from the same patron.

\begin{quote}
``Every Rune is a promise. Every line a covenant. Do not mistake the Runekeeper’s silence for weakness; their memory is the foundation of our craft.''
\end{quote}

\subsubsection{The Invokers}
\index{Invokers}

\textbf{Invokers} are those who dare. Neither archivists nor worshippers, they are travelers on the knife-edge between story and ruin. An Invoker learns the Rites of a Patron, weaves them into their own Art, and bends fate for a moment. Yet invocation is not command: it is negotiation. The Patron always leaves its mark. The stronger the Rite, the deeper the scar.

Invokers are often wanderers, exiles, or seekers. To common folk they are feared---witches, oathbreakers, meddlers with things not meant for mortal hands. But when the village falls to plague, when the sea closes its roads, when the dead refuse their rest, it is an Invoker who is called upon.

\subsubsection{Closing Words}
The dance between Patron, Runekeeper, and Invoker is a triangle of peril and necessity. Without Patrons, there is no power. Without Runekeepers, no record. Without Invokers, no action. Together, they shape the crooked, perilous art we call Invocation.

\section{Patrons \& Rites}
\label{sec:patrons-rites}


% Optional: standard macro to format each Patron’s Gift call-out the same way
\newcommand{\PatronGift}[2]{% #1 = Thematic skill, #2 = brief domain blurb
\paragraph{Patron's Gift (Imbuement).}
Once per scene as an action (cost: 1 Boon; requires \textbf{Thiasos}), touch an item to imbue it until scene end with \textbf{+1 Melee} and \textbf{+1 #1}. \emph{Push It:} extend one more scene by marking \textbf{+1 Obligation}. Gifts from the same Patron don’t stack; take the best. Dice bonuses respect the +3 cap. \textit{Domain:} #2.
}

\begin{table}[h]
\centering
\caption{Universal Push It Costs}
\begin{tabular}{|l|l|}
\hline
\textbf{Cost Component} & \textbf{Effect} \\ 
\hline
+1 SB & Escalate effect immediately \\ 
+1 Fatigue & Immediate physical/mental strain \\ 
+1 Corruption Clock Segment & Long-term Patron influence (unless otherwise specified) \\ 
GM spends 1 SB & Thematic complication (unless otherwise specified) \\ 
\hline
\end{tabular}
\end{table}

Note: Some talents, Rites, or magical paths may specify alternative corruption costs or additional consequences for Push It actions. When explicitly stated, those specific rules override the universal costs.

\paragraph{Clearing Corruption}
Corruption may be reduced through \textit{purging rituals}, such as exorcisms, sacred songs, or rites of contrition. 
These require a test (typically \textbf{Lore + Spirit}) against a DV equal to the character’s current corruption level.  
On success, reduce corruption by 1. On failure, the corruption manifests violently, imposing a temporary Condition or advancing its narrative expression.  

Optional: A \textbf{Story Beat} may also be spent to attempt such a ritual, representing the personal cost of atonement. Patrons may demand specific acts of service, sacrifice, or obligation as part of the purging process.

% Optional: standard macro to format each Patron’s Gift call-out the same way
\newcommand{\PatronGift}[2]{% #1 = Thematic skill, #2 = brief domain blurb
\paragraph{Patron's Gift (Imbuement).}
\textbf{Cost: Free, requires a Thaisos}

Once per scene as an action (cost: 1 Boon; requires \textbf{Thiasos}), touch an item to imbue it until scene end with \textbf{+1 Melee} and \textbf{+1 #1}. \emph{Push It:} extend one more scene by marking \textbf{+1 Obligation}. Gifts from the same Patron don’t stack; take the best. Dice bonuses respect the +3 cap. \textit{Domain:} #2.
}
Here is the updated table with Xhak'Thul, the Thunder-Speaker added:

\begin{table}[H]
  \centering
  \renewcommand{\arraystretch}{1.15}
  \begin{tabular}{@{}p{3.8cm}p{3.8cm}p{7.5cm}@{}}
  \toprule
  \textbf{Patron} & \textbf{+1 Thematic Skill} & \textbf{Gift / Lore Bestowal} \\
  \midrule
  Malachai (Curses, Corruption) & Command & Grants the shadow in salvation, a gift that devours from within. \\
  Aveh (Freedom, Erasure) & Stealth & Grants the storm’s passage: the gift of vanishing paths, freedom unbound by law or memory, but at the cost of leaving nothing behind. \\
  Carrion King (Carrion, Renewal) & Survival & Grants the feast of decay, where what is dead becomes seed for what lives. \\
  Clockwork Monad (Mechanism, Process) & Tinker & Grants the certainty of repetition: a cycle that never falters, a gear that never slips. \\
  Gallows Bell (Doom, Last Rites) & Command & Grants the toll of ending, a voice that closes stories and calls debts due. \\
  Ikasha (Shadow, Penumbra) & Stealth & Grants the hush between footsteps and the raven's omen at every threshold. \\
  Inaea (Mercy, Hearth) & Persuade & Grants the hearth's warmth, shelter to the weary and mercy for the lost. \\
  Inquisitor Prime (Purity, Persecution) & Insight & Grants the hunter's sight, piercing illusions and revealing supernatural corruption. \\
  Isoka (Serpents, Shedding) & Skirmish & Grants the serpent's coil, strength in sudden strike and wisdom in renewal. \\
  Khemesh (Abyss, Pressure) & Skirmish & Grants the crushing silence of the deep, where strength is drowned in weight. \\
  Livaea (Temptation, Desire) & Persuade & Grants the lure of longing, beauty sharpened into power over hearts and alliances. \\
  Lunara (Reflection, Hidden Truths) & Notice & Grants the silver quiet, finding seams and reflections that reveal what is concealed. \\
  Mab (Glamour, Courts) & Persuade & Grants the mask of favor, a voice that bends courtiers and kindles desire. \\
  Maelstraeus (Infernal Bargainer) & Persuade & Grants the contract's weight, every deal sealed in fire and shadow. \\
  Morag the Hag (Bargains, Transformation) & Deception & Grants the cunning of deals, seeing hidden costs and twisting fate through negotiation. \\
  Mor'iraath (Entropy, Destruction) & Skirmish & Grants the entropic touch, seeing weakness and consuming value in fire and ruin. \\
  Mykkiel (Zeal, Flame, Judgment) & Command & Grants the blazing verdict, a covenant carried in fire and oath, binding law to blade. \\
  Nidhoggr (Deep Earth, Rot) & Skirmish & Grants the weight of ages, the strength of stone and the hunger of roots. \\
  Oath of Flame \& Light (Dawn, Vows) & Command & Grants the fire of dawn, a vow that shields the faithful and sears the faithless. \\
  Old Man of the Black Forest (Primal Humanity, Instinct) & Survival & Grants the wild memory: fang, fire, and the path of instinct through the dark wood. \\
  Pale Shepherd (Thresholds, Guidance) & Stealth & Grants the shepherd's whisper, guiding through liminal spaces and comforting the dying. \\
  Palinode (Performance, Rapture) & Performance & Grants the breath of inspiration, turning craft into captivating art that binds audiences. \\
  Raéyn (Storm, Tides) & Skirmish & Grants the sailor's fortune: winds that shift, storms that answer to will. \\
  Sacred Geometry (Perfect Forms) & Tinker & Grants the compass of perfection, every shape reduced to its true measure. \\
  \textbf{Silent Choir (Mercy, Silence)} & \textbf{Insight} & Grants the mercy of withheld truth, the stillness that spares and erases, leaving peace where confession might have burned. \\
  \textbf{Confessor Beneath the Bell (Burden, Exchange, Guilt)} & \textbf{Command} & Grants the weight of absolution, the sacred burden that moves from sinner to witness, binding both in shared silence until the bell tolls. \\
  Sealed Gate (Boundaries, Closure) & Tinker & Grants mastery of thresholds—doors that yield or bar at your command. \\
  The Ninth Rim (Void, Negation) & Stealth & Grants mastery over absence, the power found in silence and erasure. \\
  The Traveler (Ways, Roads) & Notice & Grants the open way, a compass that never rests, and roads where none are marked. \\
  The Witness (Truth, Revelation) & Notice & Grants the unblinking gaze that unmasks deceit and remembers every oath. \\
  Thrysos (Revels, Ecstasy) & Performance & Grants the satyr’s crown, frenzy that drowns sorrow in wine and summons freedom in abandon. \\
  Varnek Karn (Ossuary, Dominion of the Dead) & Command & Grants the silence of the archive, where the dead obey and records speak. \\
  Vorthak (Consumption, Void) & Stealth & Grants the endless hunger, devouring substance and meaning alike. \\
  Xhak'Thul (Storm, Shock, Awe) & Command & Grants the voice that splits mountains, the drum that shakes foundations, power drawn from the primal storm. \\
  \bottomrule
  \end{tabular}
  \caption{Patron's Gift: fixed Thematic Skill and lore of their bestowed blessing. Thematic bonuses apply only when the fiction matches the Patron's domain.}
  \label{tab:gift-thematic-map}
\end{table}

%========================================
% appendix/patron-rivalries-expanded.tex
%========================================
% Include with: \input{appendix/patron-rivalries-expanded.tex}
% Requires: \usepackage{booktabs,longtable}

\section*{Patron Rivalries}
\label{app:patron-rivalries}

Use this matrix to quickly shade rulings. “Edge Loci” are environments or situations where one side tends to start a step better in Position or gains an Effect nudge (Keeper’s call). “Friction” are handy prompts for SB spends.

\renewcommand{\arraystretch}{1.15}
\setlength{\LTpre}{0pt}
\setlength{\LTpost}{0pt}

\begin{longtable}{@{}p{3.3cm}p{3.3cm}p{4.6cm}p{7.2cm}@{}}
  \toprule
  \textbf{Patron} & \textbf{Rival} & \textbf{Edge Loci} & \textbf{Friction \& Prompts (SB)} \\
  \midrule
  \endfirsthead

  \toprule
  \textbf{Patron} & \textbf{Rival} & \textbf{Edge Loci} & \textbf{Friction \& Prompts (SB)} \\
  \midrule
  \endhead

  \midrule
  \multicolumn{4}{r}{{Continued on next page}} \\
  \endfoot

  \bottomrule
  \endlastfoot

  Aveh (Freedom, Erasure) & Oath of Light \& Flame (Dawn, Vows) &  
  Crossroads, abandoned gates, wind-scoured battlefields. &  
  SB: vows dissolve in the storm; liberty shatters oaths; the dawn sears absence into memory. \\

  Aveh (Freedom, Erasure) & Varnek Karn (Necromantic Archives) &  
  Forgotten cairns, erased ledgers, storm-blown libraries. &  
  SB: records fade into silence; the storm devours testimony; archive cannot hold the exiled. \\

  Aveh (Freedom, Erasure) & Sealed Gate (Boundaries, Closure) &  
  Crumbling thresholds, broken wards, breached tombs. &  
  SB: the Rider tramples seals; closure collapses to exile; freedom births chaos beyond boundary. \\

  Aveh (Freedom, Erasure) & Carrion King (Decay, Renewal) &  
  Open steppes, plague winds, unmarked graves. &  
  SB: storm scatters the feast; renewal has no soil to root; liberty dissolves the cycle into absence. \\

  Aveh (Freedom, Erasure) & Gallows Bell (Justice, Judgment) &  
  Storm-dark tribunals, deserted execution grounds. &  
  SB: judgment unheard in storm’s roar; rope snaps in gale; freedom mocks justice, leaving silence. \\

  Carrion King (Carrion, Renewal) & Varnek Karn (Necromantic Archives) &
  Catacombs, bone-fields, ossuary halls. &
  SB: the feast of decay disputes record; skulls reorder themselves; scavengers mock the archive. \\

  Clockwork Monad (Process, Iteration) & The Traveler (Ways, Roads) &
  Caravans, milestones, repeating relays. &
  SB: schedule collapses; detour offers safety; routine blinds to ambush. \\

  The Confessor Beneath the Bell (Burden, Exchange, Guilt) & Oath of Light \& Flame (Dawn, Vows) &
  Burnt confessionals, sunrise absolutions, bells that toll at false dawn. &
  SB: vows cleanse through pain; guilt burns but leaves ash; the penitent's light blinds the absolver. \\

The Confessor Beneath the Bell (Burden, Exchange, Guilt) & The Witness (Truth, Revelation) &
  Tribunal halls, bloodied ledgers, whispered testimonies. &
  SB: confession births revelation; truth festers in silence; guilt migrates from sinner to judge. \\

The Confessor Beneath the Bell (Burden, Exchange, Guilt) & The Silent Choir (Mercy, Silence, Oblivion) &
  Shuttered sanctuaries, wordless absolutions, drowned bells. &
  SB: mercy silences guilt before forgiveness is granted; silence absolves but leaves the soul hollow. \\

The Confessor Beneath the Bell (Burden, Exchange, Guilt) & Sealed Gate (Boundaries, Closure) &
  Locked confessionals, sealed reliquaries, hidden sins. &
  SB: penance denied; the gate holds sin too tightly; absolution festers within confinement. \\

  Gallows Bell (Justice, Judgment) & Mykkiel (Law, Writ) &
  Courts, tribunals, execution grounds. &
  SB: judgment pronounced but precedent denied; sentence demands haunting; writ echoes in the toll. \\

  Ikasha (Shadow, Penumbra) & Oath of Light \& Flame (Dawn, Vows) &
  Twilight thresholds, oath-fires, vigil sites. &
  SB: light corners a secret; vow drags shadow forth; escape costs more than silence. \\

  Ikasha (Shadow, Penumbra) & The Witness (Truth, Revelation) &
  Silent cloisters, shaded rooms, hidden alcoves. &
  SB: hush worsens Position; shadows betray your step; deferred answers arrive with weight. \\

  Inaea (Mercy, Hearth) & Livaea (Temptation, Desire) &
  Feasting halls, chambers of intimacy, hearth-feasts. &
  SB: desire strains hospitality; passion corrodes mercy; kindness is answered with demand. \\

  Inquisitor Prime (Purity, Persecution) & Morag the Hag (Bargains, Transformation) &
  Crossroads, trial fires, cursed markets. &
  SB: bargain taints purification; soul scoured clean but scarred; purity test twists to hunger. \\

  Khemesh (Abyss, Pressure) & Raéyn (Storm, Tides) &
  Lightless holds, trenches, flooded caverns. &
  SB: abyss groans; storm maps fail; pressure crushes speech; tide demands oath. \\

  Livaea (Temptation, Desire) & Maelstraeus (Infernal Bargains) &
  Masked balls, deal-feasts, salons. &
  SB: desire sweetens the bargain; pact sealed in touch not ink; temptation costs more than coin. \\

  Lunara (Reflection, Hidden Truths) & The Witness (Truth, Revelation) &
  Mirrored chambers, silver pools, moonlit halls. &
  SB: reflection shows harsher truth; seams shine; witness misses what Lunara reveals. \\

  Mab (Glamour, Courts) & Palinode (Performance, Rapture) &
  Masques, salons, theaters. &
  SB: glamour demands encore; truth offends mask; favor must be sung aloud; court erupts at song's climax. \\

  Maelstraeus (Infernal Bargains) & Oath of Light \& Flame (Dawn, Vows) &
  Oath-stones, consecrated doors, witness trees. &
  SB: vow sears loophole shut; bargain price due early; sacred contract becomes trap. \\


  Malachai (Curses, Corruption) & Oath of Light \& Flame (Dawn, Vows) &
  Desecrated shrines, dawn-lit altars, sacred vows broken. &
  SB: a vow sanctifies but inflicts pain; the flame purges yet leaves despair. \\

  Malachai (Curses, Corruption) & The Witness (Truth, Revelation) &
  Confessionals, false miracles, dark reliquaries. &
  SB: a blessing unmasks a curse; truth corrodes faith; relics mock salvation; the witness must name corruption. \\

  Morag the Hag (Bargains, Transformation) & Maelstraeus (Infernal Bargains) &
  Markets, bargain-halls, crossroads. &
  SB: deal's true cost unveiled; haste breeds ruin; balance exacts more than offered. \\

  Mor'iraath (Entropy, Destruction) & Sacred Geometry (Order, Pattern) &
  Shattered temples, collapsing bridges, ruined cities. &
  SB: pattern dissolves into static; perfect form breaks; compass needle spins; order devoured by entropy. \\

  Nidhoggr (Dreaming Antiquity) & The Witness (Truth, Revelation) &
  Ruins, ancestor shrines, truth-seeing rites. &
  SB: revelation awakens the old; myth proven or destroyed; ancestor demands speech. \\

  Oath of Light \& Flame (Dawn, Vows) & Varnek Karn (Necromantic Archives) &
  Memorial crypts, plague pits, testaments. &
  SB: bone ledger demands payment; consecration threatens archive; unfinished oath resurrects feud. \\

  Pale Shepherd (Thresholds, Guidance) & Varnek Karn (Necromantic Archives) &
  Graveyards, liminal crossroads, shrines of passage. &
  SB: shepherd offers peace, Karn demands duty; soul cannot rest; death's duty vs death's silence. \\

  Palinode (Performance, Rapture) & Sacred Geometry (Order, Pattern) &
  Concert halls, ritual theaters, metered dances. &
  SB: form clashes with feeling; perfection dulls art; audience craves chaos in pattern. \\

  Palinode (Performance, Rapture) & Thrysos (Revels, Ecstasy) &
  Festival grounds, taverns, moots. &
  SB: song feeds revel; revel swallows song; crowd erupts beyond control; inspiration burns to mania. \\

  Raéyn (Storm, Tides) & Oath of Light \& Flame (Dawn, Vows) &
  Consecrated decks, sunrise rites at sea. &
  SB: dawn sears back abyss; vow drags crew into peril; storm promises aid with debt. \\

  Sealed Gate (Boundaries, Closure) & The Ninth Rim (Void, Negation) &
  Failing wards, thinning thresholds, sealed vaults. &
  SB: door rattles; ward flickers; void slips past; seal admits what it bars. \\

  Sealed Gate (Boundaries, Closure) & The Traveler (Ways, Roads) &
  Customs houses, locks, liminal crossings. &
  SB: ledger wrong; key refuses; toll extracted; desire path collapses to law. \\

  % --- Patron Concordance: The Silent Choir pairings ---

  The Silent Choir (Mercy, Silence) & The Witness (Truth, Revelation) &
  Echoing cloisters, confessionals with sealed lips, mirrors that reflect no mouth. &
  SB: truth yearns to speak but mercy stills the tongue; revelation shatters peace; silence becomes its own confession. \\

  The Silent Choir (Mercy, Silence) & Inaea (Mercy, Hearth) &
  Shuttered hearths, mourning halls, letters never sent. &
  SB: mercy divides; the hearth forgives through warmth, the Choir through forgetting; love dies in gentleness unspoken. \\

  The Silent Choir (Mercy, Silence) & Varnek Karn (Ossuary, Dominion of the Dead) &
  Catacombs of still air, funerary archives where the names are erased. &
  SB: the Choir hums among the bones; records fall mute; memory yields to rest; death and mercy share the same hymn. \\

  The Silent Choir (Mercy, Silence) & Sacred Geometry (Perfect Forms) &
  Anechoic temples, fractal vaults that swallow sound. &
  SB: silence perfects form; entropy refines design; mercy becomes architecture; the pattern prays without words. \\

  The Silent Choir (Mercy, Silence) & Aveh (Freedom, Erasure) &
  Ruined cloisters, desert shrines scoured by wind. &
  SB: silence liberates; erasure redeems; mercy untethers the self until nothing remains but peace. \\

  Thrysos (Revels, Ecstasy) & Carrion King (Carrion, Renewal) &
  Feast halls, bonfires, festival graves. &
  SB: revel spills to carrion; death becomes dance; feast breeds seed of decay. \\

  Thrysos (Revels, Ecstasy) & Livaea (Temptation, Desire) &
  Bacchic revels, decadent salons, wine-feasts. &
  SB: revel loosens restraint; passion binds harder; temptation drunk becomes vow. \\

  The Ninth Rim (Void, Negation) & The Witness (Truth, Revelation) &
  Forgotten vaults, failing records, silent ruins. &
  SB: truth swallows itself; record fades to dust; memory refuses recall. \\

  The Traveler (Ways, Roads) & Clockwork Monad (Process, Iteration) &
  Pilgrim routes, caravans, timetables. &
  SB: routine bends to detour; detour grants swiftness but ambush; schedule cannot hold. \\

  Varnek Karn (Necromantic Archives) & Mykkiel (Judgment, Writ) &
  Audit halls, courts martial, ledger-vaults. &
  SB: writ binds restless; sentence offends precedent; archive haunts judgment. \\

  Vorthak (Consumption, Void) & Sealed Gate (Boundaries, Closure) &
  Wards, vaults, guarded thresholds. &
  SB: hunger devours lock; seal becomes meal; barrier broken by appetite. \\

  Xhak'Thul (Storm, Shock, Awe) & Oath of Flame & Light (Dawn, Vows) & Storm-lit altars, dawn-thunder battles, sites of primal oaths sworn under sky. & SB: lightning sanctifies but burns the vow-breaker; storm drowns the dawn-cry; thunder answers fire with its own judgment. \\
\end{longtable}

\paragraph{Quick Rulings.}
\begin{itemize}
  \item \textbf{Position Nudge:} In a home locus, start one step better; in a rival locus, one step worse.
  \item \textbf{Effect Shade:} Where a Patron dominates, consider an Effect bump; where opposed, consider Limited Effect unless paid for.
  \item \textbf{Symbol Interference (Invokers):} Carrying both sides’ Symbols increases narrative noise: first ritual each scene may mark +1 Obligation (Keeper’s call).
  \item \textbf{SB Color:} When spending SB in these matchups, prefer suits that fit: Hearts (social), Spades (harm/escalation), Clubs (material cost), Diamonds (numinous disturbance).
\end{itemize}

% =========================
% OBLIGATION OVERFLOW (RITES)
% =========================
\section{Obligation Overflow (Rites)}
\label{sec:obligation-overflow}

When \textbf{Obligation} is ticked past its maximum (from Rites, vows, bargains), mark \textbf{Fatigue} based on scene severity:
\begin{itemize}
  \item \textbf{Low}: +1 Fatigue
  \item \textbf{Standard}: +2 Fatigue
  \item \textbf{High}: +3 Fatigue
\end{itemize}
If this \emph{fills} the Fatigue Track, apply the \textbf{Fatigue $\rightarrow$ Harm} conversion (see \S\ref{sec:health-fatigue-harm-rev}).

% !TEX root = resource_guide_main.tex
% Extended Section: Clearing Obligation Segments

\section{Clearing Obligation Segments (Extended)}\label{sec:obligation-clearing-extended}
\index{Obligation}\index{Patrons}\index{Acts of Service}\index{Boon}\index{Focused Devotion}

Obligation measures the growing narrative weight of a Patron’s attention. It is not debt in a transactional sense, but rather the sense that a Patron has invested, noticed, and expects relevance. This section expands the SRD guidance with philosophy, examples, and optional rules.

\subsection{Core Principle}\label{subsec:obligation-principle}
Obligation is cleared through \textbf{meaningful engagement} with the Patron. The system assumes Patrons are narrative entities: they do not care about coin, but about resonance. Clearing requires a demonstration that the character is still walking their path.

\subsection{Off-Screen Clearing (Primary Method)}\label{subsec:obligation-offscreen}
\begin{itemize}
\item \textbf{Act of Service.} Clear 1 segment by performing a resonant service aligned to the Patron’s themes or goals.\index{Acts of Service}
\item \textbf{Requirement.} The player must provide an \emph{Intricate Description} of the act—detail, symbolism, and impact.
\item \textbf{Scope.} The act need not be arduous, but it must be thematically precise (e.g., weaving a shadow tapestry for Ikasha; negotiating a shrewd pact for Maelstraeus).
\item \textbf{Reward.} Successful clearing often comes with narrative color: whispers of approval, subtle omens, or a temporary advantage in Patron-facing scenes.
\end{itemize}

\subsection{On-Screen Clearing (Optional Mechanic)}\label{subsec:obligation-onscreen}
\paragraph{Talent: Focused Devotion (3–5 XP).}\index{Focused Devotion} Once per scene, spend 1 Boon to clear 1 segment of Obligation from a Patron. This cannot remove segments accrued during the same scene. It represents a desperate or ritualized gesture to stave off attention without fully satisfying it.

\subsection{Narrative Implications}\label{subsec:obligation-implications}
\begin{enumerate}
\item \textbf{Player Agency.} The player defines how service manifests, reinforcing creativity within the Patron’s thematic frame.
\item \textbf{GM Collaboration.} The GM adjudicates sufficiency: Is the service aligned? Is it intricate enough?\index{GM Guidance}
\item \textbf{Ongoing Relationship.} Clearing segments emphasizes that the Patron bond is active, not static. The relationship evolves with acts of recognition.
\item \textbf{Scene Economy.} The cap on on-screen clearing ensures the mechanic cannot erase costs just accrued. It forces the long view.
\end{enumerate}

\subsection{Worked Examples}\label{subsec:obligation-examples}
\paragraph{Example 1: Ikasha, Angel of Serpents.} A character clears a segment by shedding their old cloak, burning it, and narrating how it symbolizes leaving behind a former self. The GM rules it sufficient, noting the subtle scent of scales in the air.

\paragraph{Example 2: Maelstraeus, Infernal Bargainer.} The player narrates writing an elaborate contract that offers a small service to a minor devil. The GM agrees: the service is thematically aligned and clears 1 segment. Later, this contract may resurface as a hook.

\paragraph{Example 3: Isoka, Angel of Change.} The group weaves discarded materials into a shrine mosaic. Its intricate description convinces the GM, who clears 1 segment and hints that a future boon will resonate with this act.

\subsection{Optional Modules}\label{subsec:obligation-modules}
\paragraph{Overflow Consequences.} When Obligation fills and overflows, the Patron demands immediate service. This may be delivered as a Complication Point (or Story Beat) trigger: the Patron intrudes, offering aid with a binding cost.\index{Overflow}
\paragraph{Segment Scaling.} Particularly devoted characters may require greater services: clear 2 segments with a dramatic act (pilgrimage, sacrifice, or political maneuver), subject to GM approval.
\paragraph{Shared Obligation.} In multi-character bonds, the GM may allow a group act of service to clear multiple segments, but only if each player contributes to the Intricate Description.

\subsection{GM Guidance}\label{subsec:obligation-gm}
\begin{itemize}
\item \textbf{Be Generous with Hooks.} Reward creativity with flavorful omens, not just a cleared box.
\item \textbf{Maintain Tension.} Do not allow trivial or repetitive services to count. Require new symbolic angles each time.
\item \textbf{Keep On-Screen Costly.} Remind players that Focused Devotion costs a Boon and cannot erase fresh debt.
\item \textbf{Tie to Patron Themes.} If a service drifts off-theme, let it fail or only partially clear.
\end{itemize}

\begin{tcolorbox}[title={Summary},colback=gray!5,colframe=black]
Obligation is cleared through \textbf{thematic service, intricate description, and narrative engagement}. Off-screen acts remain the primary path; on-screen clearing is a costly safety valve. Treat every act as a chance to show how characters evolve under their Patron’s gaze.
\end{tcolorbox}

\section{Acts of Service — Examples}\label{sec:acts-of-service-table}
\index{Acts of Service}\index{Obligation}\index{Patrons}

Patrons delight in resonance, not labor. Use these examples as prompts---\emph{petty} butterflies and \emph{involved} operations both count when on-theme. Each entry lists a scale suggestion and what it might clear.
  
\subsection*{GM Notes}\label{subsec:acts-service-gm}
\begin{itemize}
\item \textbf{Novelty Over Volume.} Repeating the same petty act does not keep clearing segments; vary the symbol or site.
\item \textbf{Precision Beats Difficulty.} A sharp thematic hit at “petty” scale can clear 1 segment as surely as a muddled “modest.”
\item \textbf{Hook the World.} Many of these acts conveniently create future clocks, NPC ties, or bargains—use them.
\item \textbf{Patron’s Largess} A Patron’s will is mysterious, but their displeasure is rarely subtle. For narrative reasons, the GM may rule that a Patron is especially offended—perhaps by a neglected omen, a rival gaining favor, or simple caprice—and exact an additional toll beyond the usual cost.
\end{itemize}

\subsection{Patron Fear Influence Table}
\label{subsec:patron-fear-table}
\begin{longtable}{@{}p{3.2cm}p{4.5cm}p{2.8cm}p{5.9cm}@{}}
\toprule
\textbf{Patron} & \textbf{Domain Tags} & \textbf{Influence Type} & \textbf{Fear Effect Modifications} \\
\midrule
\endfirsthead

\multicolumn{4}{c}{{\bfseries Continued from previous page}} \\
\toprule
\textbf{Patron} & \textbf{Domain Tags} & \textbf{Influence Type} & \textbf{Fear Effect Modifications} \\
\midrule
\endhead

\midrule
\multicolumn{4}{r}{{Continued on next page}} \\
\endfoot

\bottomrule
\endlastfoot

Aveh, the Rider Behind the Storm & Freedom, Exile, Erasure & Reroll & 
Fear tagged \emph{Storm, Exile, Freedom} shifts unpredictably. On 5 (Hide), victim seems to vanish—companions may forget their presence until reminded. \\

Carrion King, Lord of Renewal & Decay, Feast, Scavenging & Reroll & 
Fear tagged \emph{Decay, Hunger, Feast} mutates. On 1 (Fight), frenzy devolves into cannibalistic hunger. \\

Clockwork Monad, Engine of Iteration & Order, Repetition, Mechanism & Amplify & 
Fear tagged \emph{Order, Pattern, Repetition} becomes mechanistic horror. On 3 (Freeze), victim is trapped in endless loop. \\

The Confessor Beneath the Bell, Bearer of Sins & Burden, Exchange, Guilt & Transfer & 
Fear tagged \emph{Burden, Guilt, Confession} does not fade when resisted—it migrates. 
On 2 (Hide), the victim’s guilt manifests in another nearby; on 4 (Flee), they feel the weight shift, forcing them to drop possessions or allies as though casting off sin. 
On 6 (Submit), they confess uncontrollably, transferring their terror and shame into the listener. \\

Gallows Bell, Doom of Endings & Doom, Last Rites & Resonance & 
Fear tagged \emph{Ending, Toll, Justice} deepens. On 2 (Hide), victim hears the bell toll and accepts inevitable demise. \\

Ikasha, She Who Sleeps & Shadow, Whispers, Latent Potential & Amplify & 
Fear tagged \emph{Shadow, Voice, Hidden} deepens. On 3 (Freeze), victim may be paralyzed by whispered secrets. \\

Inaea, Angel of the Spider & Mercy, Hearth, Connection & Reroll & 
Fear tagged \emph{Home, Family, Comfort} twists to terror of loss. On 4 (Flee), victim abandons loved ones "for their own good." \\

Inquisitor Prime, Purity's Flame & Purity, Zeal, Persecution & Amplify & 
Fear tagged \emph{Purity, Heresy, Persecution} turns fanatical. On 1 (Fight), victim attempts self-destructive martyrdom. \\

Khemesh, the Abyssal Maw & Abyss, Pressure, Depths & Amplify & 
Fear tagged \emph{Depth, Pressure, Dark} crushes. On 1 (Fight), victim dives berserk into consuming dark. \\

Livaea, Crimson Courtier & Temptation, Desire, Passion & Amplify & 
Fear tagged \emph{Desire, Passion, Temptation} becomes seductive and lethal. On 6 (Submit), victim surrenders willingly to corrupt longing. \\

Lunara, Mirror of Truths & Reflection, Mirrors, Secrets & Amplify & 
Fear tagged \emph{Reflection, Mirror, Truth} disturbs. On 3 (Freeze), victim may be trapped by their own image. \\

Mab, Queen of Courts & Glamour, Courts, Deception & Reroll & 
Fear tagged \emph{Illusion, Social, Performance} may be reshaped. On 6 (Submit), victim suffers elaborate humiliation instead of surrender. \\

Maelstraeus, Infernal Bargainer & Contracts, Deals, Fire & Resonance & 
Fear tagged \emph{Bargain, Deal, Contract} imposes Complication Points. Fear of infernal agreement has lasting cost. \\

Malachai, the Chained Angel & Curses, Corruption & Amplify & 
Fear tagged \emph{Curse, Despair, Shadow} worsens. On 3 (Freeze), victim may lapse into catatonic acceptance of corruption. \\

Morag the Hag & Transformation, Bargains, Cunning & Reroll & 
Fear tagged \emph{Transformation, Trickery, Deal} warps into faustian dread. On 5 (Hide), victim changes shape grotesquely. \\

Mor'iraath, Ashen Maw & Entropy, Destruction, Endings & Amplify & 
Fear tagged \emph{Destruction, Entropy, End} becomes apocalyptic. On 2 (Hide), victim seeks oblivion in decay. \\

Mykkiel, Arbiter of the Writ & Judgment, Law, Authority & Resonance & 
Fear tagged \emph{Authority, Judgment, Law} adds marks to Fear Track. The burden of judgment lingers. \\

Nidhoggr, World-Worm & Dream, Antiquity, Buried Truths & Resonance & 
Fear tagged \emph{Dream, Ancient, Truth} also inflicts Fatigue. Weight of aeons drains resolve. \\

Oath of Flame \& Light & Dawn, Vows, Purity & Resonance & 
Fear tagged \emph{Light, Sacred, Vow} also imposes Fatigue. Divine light turns terror into bodily strain. \\

Pale Shepherd, Guide of Thresholds & Thresholds, Souls, Passing & Reroll & 
Fear tagged \emph{Death, Guidance, Soul} becomes existential. On 6 (Submit), victim follows willingly into afterlife path. \\

Palinode, Queen of Encores & Performance, Song, Rapture & Resonance & 
Fear tagged \emph{Performance, Art, Audience} adds Complication Points. Fear of failure binds mechanically. \\

Raéyn, Mistress of the Sea & Sea, Tides, Travel & Reroll & 
Fear tagged \emph{Water, Lost, Travel} shifts unpredictably. On 5 (Hide), victim may slip into fog or sink beneath waves. \\

Sealed Gate, Keeper of Thresholds & Boundaries, Closure & Amplify & 
Fear tagged \emph{Confinement, Threshold, Trapped} worsens. On 4 (Flee), victim may injure themselves clawing at sealed barriers. \\

The Silent Choir, Mercy of the Unspoken & Mercy, Silence, Oblivion & Suppress & 
Fear tagged \emph{Silence, Guilt, Confession} is stifled rather than resisted. On 3 (Freeze), victim cannot scream or speak; on 5 (Beg), they fall to their knees, mouthing prayers that make no sound. \\

The Ninth Rim, the Unseen Ninth & Void, Negation & Reroll & 
Fear tagged \emph{Void, Absence, Silence} may be rerolled to reflect conceptual horror. On 2 (Hide), victim may vanish into existential erasure. \\

The Traveler, Road-Sworn & Ways, Journeys, Crossroads & Resonance & 
Fear tagged \emph{Lost, Path, Journey} also inflicts Complication Points. Terror of losing the way becomes mechanical hindrance. \\

The Witness, Unblinking Eye & Truth, Revelation & Amplify & 
Fear tagged \emph{Truth, Knowledge, Revelation} intensifies. On 6 (Submit), target may confess dark secrets under perceived divine judgment. \\

Varnek Karn, Bone King & Necromantic Archives, Death & Amplify & 
Fear tagged \emph{Death, Undead, Necromancy} becomes horrific. On 2 (Hide), victim feigns death among corpses. \\

Vorthak, Withered Crown & Hunger, Void, Consumption & Amplify & 
Fear tagged \emph{Hunger, Consumption, Void} devours. On 4 (Flee), victim frenzies to consume others. \\

Xhak'Thul, Thunder-Speaker & Storm, Shock, Awe, Primal Memory & Amplify & Fear tagged \emph{Storm, Thunder, Primal Terror} intensifies. On 1 (Fight), victim is overwhelmed by berserker fury. On 5 (Hide), they flee into the storm itself. \\

\end{longtable}

\subsection{Implementation Guidelines}
\label{subsec:patron-fear-guidelines}

\subsubsection{Tag Matching}
When a Fear effect is triggered, compare the effect's Magic Tags against the active Patron's Domain Tags. If any tags match, the Patron's influence applies.

\subsubsection{Multiple Patrons}
If multiple Patrons are present with overlapping domains, the GM may:
\begin{itemize}
  \item Choose the most thematically appropriate influence
  \item Apply influences sequentially (reroll then amplify)
  \item Allow players to choose which influence affects them
\end{itemize}

\subsubsection{Player Agency}
Players should be informed when Patron influence is active and given the opportunity to:
\begin{itemize}
  \item Spend resources (Boons, Willpower) to resist the influence
  \item Narrate how their character resists or embraces the effect
  \item Choose from available influence options when multiple apply
\end{itemize}

\subsection{Embracing Corruption}
\index{Corruption!Embracing} \label{subsec:embracing-corruption}

A character with \textbf{2+ levels of Corruption} may choose to \textbf{Embrace} it, transforming creeping decay into a permanent \textbf{Talent}. Embracing never reduces Corruption --- it reshapes it. The deeper the corruption, the greater the power and the cost.

\begin{center}
\begin{tabular}{@{}l l l@{}}
\toprule
\textbf{Corruption Level} & \textbf{Talent Tier} & \textbf{Cost} \\
\midrule
2 (Early) & Minor Talent & +1 Permanent Corruption (cannot Fade below 1) \\
3 (Deepening) & Major Talent & +2 Permanent Corruption, constant visible tell \\
4+ (Severe) & Transcendent Talent & Permanently altered, locked at Stage 3+ \\
\bottomrule
\end{tabular}
\end{center}

\noindent Embraced Talents always reflect the Patron’s themes (e.g., shadow, indulgence, empathy, paranoia). They grant great power, but mark the character irrevocably as claimed.