% !TEX root = ../fates_edge_players_guide.tex

\chapter{Gods, Powers, and Patrons}
\label{ch:gods-powers-patrons}

\begin{multicols}{2}

\index{gods}\index{powers}\index{patrons}\index{rites}\index{obligation}
In \textit{Fate’s Edge}, the \textbf{divine} is not a distant thesis—it is an active pressure system that pushes back when named. Temples argue in \emph{Ecktic High} (\S\ref{ch:languages}), bells speak law in the Mistlands, and oaths sworn under clear sky bind tighter than iron (\S\ref{ch:world-regions}). This chapter restores the \textbf{lore of traditions}, clarifies \textbf{Patrons} and their \textbf{Rites}, and shows how all of it interlocks with the core systems (\S\ref{ch:magic}).

\section{The Nature of Divine Forces}
\index{divine forces}\index{numina}\index{offices (divine)}
Scholars of Theona write that divinity condenses in three ways:
\begin{description}
  \item[\textbf{Numina}] \index{numina} Place-bound or concept-focused powers (a harbor’s luck, a winter’s severity).
  \item[\textbf{Lineages}] \index{lineages!divine} Ancestor-chains and civic cults that accrue obligation over centuries.
  \item[\textbf{Offices}] \index{offices (divine)!seats} Roles that exist whether or not a face is seated—\emph{Witness}, \emph{Arbiter}, \emph{Wayfinder}. Whoever answers the call \emph{is} the office for as long as the rites hold.
\end{description}
All three \emph{answer} when approached in their proper language, place, and price (\S\ref{ch:languages}).

\paragraph{Friction With Reality.}\index{backlash}\index{setback points (SP)}
Calling on the sacred generates narrative pressure. Treat divine missteps as Setback Points (SP) that the GM can spend for omens, visitations, taboos invoked, or social consequences among the faithful (\S\ref{ch:magic}).

\section{Major Divine Traditions}
\index{divine traditions}\index{cults}\index{temples}

\subsection*{The Everflame and the Lampers}
\index{Everflame}\index{Lampers@Lampers (urban confraternities)}
\textbf{Domains:} Fire, purification, law \quad
\textbf{Centers:} Ecktoria’s marble courts, Sun Coast porticoes \\
\textbf{Themes:} Order through light, confession by heat, civic duty
\begin{itemize}
  \item \textbf{Clergy \& Orders:} \emph{Lampers} tend beacon-fires and court lamps; \emph{Ash-Deacons} oversee penances. \index{orders!Lampers}\index{orders!Ash-Deacons}
  \item \textbf{Rites:} \emph{Lamp of Plain Speech} (forces truth in sworn testimony), \emph{Cinder-Ward} (burns contagion, literal or social).
  \item \textbf{Customs:} \emph{Three-Wick Vigil} before verdicts; extinguishing a lamp early is a public insult. \index{customs!three-wick vigil}
\end{itemize}
\emph{Vignette:} \textit{In the Echo Hall a thousand glass panes bloom with flame. Each verdict adds one. The Archivist snuffs two and the crowd inhales—history itself has been corrected.}

\subsection*{The Reformed Light (Mercy-Cant)}
\index{Light Reformation}\index{mercy-cant}
\textbf{Domains:} Revelation, mercy, community \quad
\textbf{Centers:} Valewood clinics, Theonan schools \\
\textbf{Themes:} Repair over punishment; unity through service
\begin{itemize}
  \item \textbf{Orders:} \emph{Lantern-Hands} (healers), \emph{Keepers of Account} (community ledgers of generosity). \index{orders!Lantern-Hands}\index{orders!Keepers of Account}
  \item \textbf{Rites:} \emph{Sharing-Bread} (multiplies supplies within a bounded fellowship), \emph{Pardon-Mark} (removes a social stigma clock if restitution is made).
  \item \textbf{Taboo:} No light is to be used to shame the poor in public.
\end{itemize}

\subsection*{Sea and Storm Courts}
\index{Sea and Storm}\index{courts!maritime}
\textbf{Domains:} Oceans, storms, navigation \quad
\textbf{Centers:} Linn and Zakov archipelagos (\S\ref{ch:world-regions}) \\
\textbf{Themes:} Bargain with weather; respect for hazard
\begin{itemize}
  \item \textbf{Priesthood:} \emph{Tide-Readers} interpret current-scripts; \emph{Bell-Helms} lead weather-rites. \index{orders!Tide-Readers}\index{orders!Bell-Helms}
  \item \textbf{Rites:} \emph{Stern-Knot} (locks a course against drift), \emph{Storm’s Portion} (you surrender cargo; the squall spares lives).
  \item \textbf{Language:} Sea-Patter requires triadic call-and-response (\S\ref{ch:languages}); missing the third call risks an omen. \index{Sea-Patter}
\end{itemize}

\subsection*{Sky and Conflict Traditions}
\index{Sky}\index{Conflict}
\textbf{Domains:} Sky, conflict, honor \quad
\textbf{Centers:} Vilikari highlands, Acasian Marches \\
\textbf{Themes:} Oaths make war bearable; victory is \emph{clean} when witnessed
\begin{itemize}
  \item \textbf{Oath-Forms:} Horn-codes and duel formulae (\S\ref{ch:languages}) bind reparations. \index{horn-codes}
  \item \textbf{Rites:} \emph{Sky-Witness} (freezes a parley one exchange; no one may strike), \emph{Banner-Balance} (equalizes a skirmish’s scale for one beat).
\end{itemize}

\subsection*{Shadow and Transformation Paths}
\index{Shadow}\index{Transformation}
\textbf{Domains:} Shade, secrets, change \quad
\textbf{Centers:} Eastern caravan cloisters, Mistland bell precincts \\
\textbf{Themes:} Hidden knowledge; necessary metamorphosis
\begin{itemize}
  \item \textbf{Initiates:} \emph{Veil-Bearers} trade names seasonally (see Fog-Names, \S\ref{ch:languages}). \index{Veil-Bearers}\index{Fog-Names}
  \item \textbf{Rites:} \emph{Veil of Change} (mask a social role until challenged by bell or oath), \emph{Moth-Path} (take a consequence now to step past a danger later).
\end{itemize}

\subsection*{Ancestral Houses and Stone-Law}
\index{Ancestral veneration}\index{Stone-Law}
\textbf{Domains:} Kin, craft, memory \quad
\textbf{Centers:} Stone Kingdoms, Aeler tunnels \\
\textbf{Themes:} Work remembered; promises become architecture
\begin{itemize}
  \item \textbf{Practices:} Charter-Script read aloud to attest memory (\S\ref{ch:languages}). \index{Charter-Script}
  \item \textbf{Rites:} \emph{Line-Recall} (perfect recall in spaces your kin laid), \emph{Load-Share} (shift harm from structure to sworn caretakers as Fatigue).
\end{itemize}

\section{Patrons and the Rites Economy}
\index{patrons}\index{rites}\index{obligation}\index{attention}
Not all power wants worship. \textbf{Patrons} are entities or \emph{offices} that answer to named procedures and ledgered \textbf{Obligation}. Some wear faces. Others are the role itself.

\subsection*{The Nature of Patron Bargains}
\begin{itemize}
  \item \textbf{Rites over Prayer:} Name the Rite, prepare the token, speak the register (\S\ref{ch:languages}). No hymns; just contracts. \index{contracts, divine}
  \item \textbf{Obligation Tracks:} Each Patron has a 9-segment track. Mark segments when you invoke a Rite; clear by service, offering, or propagation (see below).
  \item \textbf{Attention:} High Obligation invites unsolicited interventions—boons or demands—framed as GM SP spends.
\end{itemize}

\subsection*{Common Patron Offices (Archetypes)}
\index{Witness (office)}\index{Wayfinder (office)}\index{Keeper of Boundaries}\index{Pattern-Weaver}\index{Catalyst}\index{Arbiter}
\begin{description}
  \item[\textbf{The Witness}] Truth, record, accountability. \emph{Rite:} \emph{Lamp of Plain Speech}. \emph{Service:} Submit copies to a Great Library (\S\ref{ch:compendium-assets}).
  \item[\textbf{Keeper of Boundaries}] Limits, wards, thresholds. \emph{Rite:} \emph{Corner-Bind} (draws a chalk line that takes a beat to cross unless named). \emph{Service:} Maintain waystones. 
  \item[\textbf{Wayfinder}] Paths, crossings, rescue. \emph{Rite:} \emph{Third Turning} (once per journey, reroute around a blocked route at a cost). \emph{Service:} Light cairns, teach safe passages.
  \item[\textbf{Pattern-Weaver}] Structure, craft, interconnection. \emph{Rite:} \emph{True Grain} (spot the hidden fault in a plan or timber). \emph{Service:} Repair what you reveal.
  \item[\textbf{Catalyst}] Change, metamorphosis. \emph{Rite:} \emph{Shed-Skin} (move a Condition from self to scene). \emph{Service:} See change through.
  \item[\textbf{Arbiter}] Weighing, recompense. \emph{Rite:} \emph{Balance Mark} (binds both sides to a next step in restitution). \emph{Service:} Enforce the step impartially.
\end{description}

\subsection*{Obligation Levels (Guidance)}
\index{obligation!levels}
\begin{center}
\small
\begin{tabular}{ll}
\toprule
\textbf{Segments} & \textbf{Typical Consequences} \\
\midrule
1--2 & Minor omens; requests by proxy; small taboos surface \\
3--5 & Regular demands; rival attention; gifts tied to errands \\
6--8 & Direct visitations; service deadlines; boons with barbs \\
9+  & Dominant influence; loss of choice without costly defiance \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Clearing Obligation.}\index{obligation!clearing}
\emph{Service} (perform themed tasks), \emph{Offerings} (sacrifice time, goods, or status), \emph{Propagation} (teach the Rite, spread the sign), or \emph{Boons} (spend advancement to buy down 1--2 segments).

\section{Engagement Paths}
\index{devotion}\index{thaumaturgy}\index{pact magic}

\subsection*{Devotional Practice}
\index{devotion!practice}
\emph{Prayer, calendar rites, ethics.} Grants access to themed \textbf{Talents} (e.g., \emph{Purifying Touch}, \emph{Guide’s Wisdom}) at normal XP cost (\S\ref{ch:talents}). Use language, posture, and season for fictional advantage (\S\ref{ch:languages}).

\subsection*{Thaumaturgical Access}
\index{thaumaturgy}
Treat miracles as procedures: preconditions, components, and test points. Use \emph{Wits + Arcana} to \emph{Channel}, then \emph{Weave} (\S\ref{ch:magic}); divine sources may swap in \emph{Spirit + Faith} for either phase, per tradition.

\subsection*{Pact Magic (Rites)}
\index{pact magic}\index{rites!invocation}
\begin{itemize}
  \item \textbf{Prereqs:} Talent such as \emph{Familiar Bond} or \emph{Codex of Names}. \index{talents!Familiar Bond}\index{talents!Codex of Names}
  \item \textbf{Invoke:} Spend a \emph{Boon} \emph{or} mark \emph{Obligation} (GM sets segments based on scale: minor 1, standard 2, major 3+).
  \item \textbf{Risk:} On misses, GM spends SP as \emph{Backlash} flavored by the Patron’s office (\S\ref{ch:magic}).
\end{itemize}

\section{Regional Faces \& Sacred Geography}
\index{regional religion}\index{sacred sites}\index{pilgrimage}

\subsection*{Inland Sea Circuit}
\index{Inland Sea}
\textbf{Pilgrim’s Ring}—a route of beacon-towers, courts, and granaries. Completing the ring in a year and a day grants the title \emph{Road-Kin} (merchants will open a door once for free). \index{Road-Kin}

\subsection*{Mistlands Bell-Precincts}
\index{Mistlands!bells}\index{Bell-Speech}
Bells carry law and lament. The \emph{Three-Tone Dusk} is binding: certain Rites (\emph{Veil of Change}) must answer or disperse harmlessly.

\subsection*{Stone Kingdoms Oath-Works}
\index{Stone Kingdoms}\index{Stone-Speech}
Bridges and vaults are “signed” with ancestor marks. A \emph{Line-Recall} Rite in the right hall reads generations like a book.

\subsection*{Western Islanders’ Weather-Courts}
\index{Western Islanders}\index{Sea courts}
Courts convene on the sea itself: three boats, three judges, verdict by current and chant (\S\ref{ch:languages} Sea-Patter).

\section{Sample Rites (Playable Seeds)}
\index{rites!examples}\index{mechanics}

\paragraph{Lamp of Plain Speech (Witness).}
\begin{itemize}
  \item \emph{Setup:} Three-wick lamp, names in Ecktic High.
  \item \emph{Roll:} \textbf{Spirit + Presence} vs DV set by audience status.
  \item \emph{Effect:} For this exchange, falsehoods spoken under oath ring dull to all present; gain a \emph{Boon} to detect contradictions.
  \item \emph{Obligation:} +1 if used for private gain; +2 if used to humiliate publicly.
\end{itemize}

\paragraph{Corner-Bind (Keeper of Boundaries).}
\begin{itemize}
  \item \emph{Setup:} Chalk four corners, speak a boundary-name from Charter-Script.
  \item \emph{Roll:} \textbf{Wits + Lore} vs DV 2–4 (site sanctity).
  \item \emph{Effect:} Crossing the threshold takes one beat of declared effort unless the crosser’s name is spoken from within.
  \item \emph{Backlash:} GM may mark \emph{“You Named Me”} on a hostile if you used their true name improperly.
\end{itemize}

\paragraph{Third Turning (Wayfinder).}
\begin{itemize}
  \item \emph{Setup:} Spin a coin at a three-way and whistle Wind-Speech. \index{Wind-Speech}
  \item \emph{Roll:} \textbf{Wits + Survival} vs weather DV.
  \item \emph{Effect:} Reveal a safe detour; costs +1 day or a tangible trade-off.
  \item \emph{Obligation:} +1 unless you cairn the path for others.
\end{itemize}

\paragraph{Shed-Skin (Catalyst).}
\begin{itemize}
  \item \emph{Setup:} Burn an old token of self.
  \item \emph{Roll:} \textbf{Spirit + Resolve} vs DV of the looming danger.
  \item \emph{Effect:} Move one \emph{Condition} from you to the \emph{Scene} as a fleeting hazard you can then exploit or endure.
  \item \emph{Complication:} A known tie reacts poorly; GM marks a \emph{Social Drift} clock.
\end{itemize}

\section{Talents \& Devotional Boons}
\index{talents!divine}\index{boons!divine}
Examples that fit the above traditions (costs per \S\ref{ch:talents}):
\begin{itemize}
  \item \textbf{Purifying Touch}—Once per scene, cleanse a mundane taint or calm a crowd segment if you carry fire openly.
  \item \textbf{Guide’s Wisdom}—When mediating, ask the GM one clarifying question about \emph{stakes}; answer is candid.
  \item \textbf{Storm Sense}—You always know the fastest safe anchorage within a day’s sail.
  \item \textbf{Oath-Keeper}—When you publicly accept terms, gain a \emph{Boon} to the next roll that advances them.
\end{itemize}

\section{Consequences \& Conflicts}
\index{consequences!divine}\index{conflict!religious}
\begin{itemize}
  \item \textbf{Attention:} High Obligation grants visitations—aid that arrives \emph{with} a demand.
  \item \textbf{Rival Doctrines:} Lampers vs Mercy-Cant on punishment; Sea Courts vs Sky Oath on jurisdiction; Ancestral Houses vs Catalyst on continuity.
  \item \textbf{Blasphemy as Mechanics:} GM spends SP to flip a rite’s tone (e.g., \emph{Lamp of Plain Speech} exposes \emph{your} omission first).
\end{itemize}

\section{GM Tools}
\index{GM tools!divine}\index{oracles!religious}
\subsection*{Quick SP Menu (Divine)}
\begin{itemize}
  \item \textbf{1 SP}: An omen complicates timing or tone (bell toll, lamp gutter).
  \item \textbf{2 SP}: A taboo is invoked; someone important is offended.
  \item \textbf{3 SP}: A rival office answers as well (Keeper \emph{vs} Catalyst).
  \item \textbf{4 SP}: An owed service comes due \emph{now}.
\end{itemize}

\subsection*{Sacred Site Tags}
\index{sacred sites!tags}
\textit{Consecrated}, \textit{Witnessed}, \textit{Thresholded}, \textit{Storm-Favored}, \textit{Ancestral-Resonant}. Each tag lowers DV by 1 for aligned rites and raises it by 1 for opposed offices.

\begin{tcolorbox}[colback=purple!5!white,colframe=purple!75!black,title=Divine \& Patron Quick Reference,fonttitle=\bfseries]
\textbf{Approaches:} Devotion (Talents), Thaumaturgy (procedures), Pact (Rites + Obligation) \\
\textbf{Obligation Track:} 9 segments—at 3/6/9 trigger requests, visitations, demands \\
\textbf{Backlash:} GM SP spend, flavored by office and region \\
\textbf{Language Keys:} Ecktic High (oaths), Sea-Patter (weather), Charter-Script (memory) \\
\textbf{Assets Help:} Libraries, colleges, shrines modify DV/Position (\S\ref{ch:compendium-assets})
\end{tcolorbox}

\section{Mechanical Integration}
\index{mechanics!divine}\index{rites!mechanics}
\begin{itemize}
  \item \textbf{Talent Access}—Divine-themed Talents use normal XP (\S\ref{ch:talents}); prerequisites are narrative.
  \item \textbf{Rites System}—Use \emph{Obligation} and SP as described; scaling: minor (1 seg), standard (2), major (3+).
  \item \textbf{Casting Loop}—Rites can slot into \emph{Channel/Weave/Backlash} (\S\ref{ch:magic}); \emph{Spirit} often substitutes for \emph{Wits}.
  \item \textbf{Language Hooks}—Correct register can lower DV by 1 or grant \emph{Boon} (\S\ref{ch:languages}).
\end{itemize}

\section*{Closing Note}
\index{allegiance}\index{story hooks}
Every allegiance writes \emph{terms} into your story: who will answer when you call, and who will come to collect when you don’t.

\end{multicols}
