\chapter{Compendium of Assets}
\label{ch:compendium-assets}

\index{assets}\index{holdings}\index{influence}
In \textit{Fate's Edge}, \textbf{Assets} are your off-screen levers—properties, charters, networks, and institutions that change the fiction between scenes. They do not fight beside you (see \S\ref{ch:assets-followers} for \emph{Followers}) but they shape the field you enter.

\section{Design Principles}
\index{assets!design principles}
Assets are written to:
\begin{itemize}
  \item \textbf{Shift fiction} between scenes: open doors, quiet rumors, stock larders, calm crowds.
  \item \textbf{Create hooks}: each Asset suggests opportunities \emph{and} complications if neglected.
  \item \textbf{Stay narrative-first}: prefer position, scale, and access over flat modifiers.
\end{itemize}

Assets do \textbf{not}:
\begin{itemize}
  \item Act inside a scene on their own (Followers do that, \S\ref{ch:assets-followers}). \index{followers|see{assets and followers}}
  \item Print wealth or XP automatically.
  \item Stack identical tags—if two Assets would grant the same edge, take the best one. \index{assets!stacking}
\end{itemize}

\paragraph{Narrative First, Mechanics Second}
When an Asset grants “start \textbf{Dominant}”, it applies to social or informational rolls where positioning matters. For purely practical challenges (e.g., picking a lock, repairing a device), “reduce DV by 1” is the preferred phrasing. When in doubt, the GM chooses the form that best fits the fiction.

\section{Asset Mechanics (Summary)}
\index{assets!mechanics}\index{upkeep}\index{Story Beats (SB)}
All Assets share the following properties (expanded rules: \S\ref{ch:assets-followers}):
\begin{description}
  \item[Cost] Minor \textbf{4 XP}, Standard \textbf{8 XP}, Major \textbf{12 XP}. \index{assets!minor}\index{assets!standard}\index{assets!major}
  \item[Establishment] 1 day / 1 week / 1 month respectively to become operational.
  \item[Upkeep] Each downtime, either pay XP equal to the Asset's base cost \emph{or} devote a full scene tending it. Neglect degrades condition (Maintained $\rightarrow$ Neglected $\rightarrow$ Compromised). \index{assets!conditions}\index{assets!maintenance}
  \item[Free Use] Once per session, trigger the listed \emph{off-screen} benefit (framed at scene start or as flashback with GM assent).
  \item[Scene Surge] Spend \textbf{1 Boon} to turn an Asset into a dramatic \emph{on-screen} intervention as described (one beat, fiction permitting). \index{Boons}
  \item[Extra Use] Spend \textbf{2 XP} to trigger the off-screen benefit an additional time this session.
  \item[Risk Interface] When an Asset covers risk, it typically \textbf{improves Position}, \textbf{lowers DV by 1}, \textbf{fills 1 segment on a relevant clock}, or \textbf{mitigates 1 SB} (Story Beat) from a thematically linked complication. \index{position}\index{difficulty value (DV)}\index{clocks}\index{Story Beats (SB)}
\end{description}

\paragraph{Capacity (Optional Rule)}\index{assets!capacity}
A character comfortably maintains a number of Assets up to their \textbf{Spirit} rating. For each Asset beyond this, increase its upkeep by +50\% (round up) \emph{or} require an additional downtime scene to keep it \emph{Maintained}.

\emph{Example: Spirit 3 → 3 Assets are easy; a 4th Asset costs 50\% more upkeep (e.g., 6 XP instead of 4) each downtime, or demands a second personal scene to maintain.}

\paragraph{Cross‑Reference to Obligation}
Assets tied to a specific faction (e.g., Temple Complex, Banking Charter, Guild Seat) may also generate \textbf{Obligation} if used aggressively or neglected. See \S\ref{ch:world-powers-obligation} for terrestrial Patron mechanics.

\section{Minor Assets (4 XP)}
\index{assets!minor}
Small holdings, charters, or kits that grant reliable leverage.
\begin{description}
  \item[\textbf{Signature Weapon}] \index{signature weapon} \emph{A named heirloom or masterpiece.} \emph{Free Use:} If gear damage would compromise it, downgrade to a minor flaw and clear with a quiet maintenance scene. \emph{Surge:} Produce it at the perfect moment; improve Position one step for your first attack this scene.

  \item[\textbf{Scholar's Cell}] \index{scholar's cell} \emph{A quiet room, permits, and a custodian.} \emph{Free Use:} Begin one research scene at \textbf{Dominant} (or lower DV by 1). \emph{Surge:} Reveal a cited source now; fill 1 segment on a relevant \emph{Information} clock.

  \item[\textbf{Safehouse Network (Quarter)}] \index{safehouse} \emph{Trapdoors, false walls, and a landlord who never saw you.} \emph{Free Use:} Start an entry/exit scene \textbf{Dominant}. \emph{Surge:} Produce a hidden egress; convert one \emph{pursuit} consequence into a temporary complication you can interact with.

  \item[\textbf{Artisan's Bench}] \index{artisan's bench} \emph{Tools for field repairs and fabrication.} \emph{Free Use:} Remove \emph{Compromised} from one mundane item between scenes. \emph{Surge:} Improvise a single-use tool; gain a \emph{Boon} on the immediate Craft check.

  \item[\textbf{Courier Chain}] \index{courier network} \emph{Runners, dead-drops, and codes in one city.} \emph{Free Use:} Move a message or small parcel safely; start the related parley \textbf{Dominant}. \emph{Surge:} A courier appears now; mitigate 1 SB tied to a missed appointment or lost message.

  \item[\textbf{Small Holding (Farm/Orchard)}] \index{farm}\index{orchard} \emph{Steady larder and a trusted foreman.} \emph{Free Use:} Waive a minor Supply complication or clear Fatigue 1 for one PC during downtime. \emph{Surge:} Produce a timely wagon, mule, or crates to justify a plausible cover story.

  \item[\textbf{Neighborhood Inn/Tavern}] \index{inn}\index{tavern} \emph{A familiar house of bread and rumor.} \emph{Free Use:} First social opener in that neighborhood starts \textbf{Dominant}. \emph{Surge:} Call in the proprietor; shift a hostile crowd one step toward neutral.

  \item[\textbf{Ritual Niche/Shrine}] \index{shrine}\index{ritual niche} \emph{A consecrated corner with a small flame.} \emph{Free Use:} Clear Fatigue 1 or mitigate 1 SB from spiritual or psychological complications (fear, despair, doubt). \emph{Surge:} Sanctify a doorway for one scene; hostile supernatural presence must reveal itself or back away one band (\S\ref{ch:world-interaction}).

  \item[\textbf{River Skiff/Barge Share}] \index{boat}\index{barge} \emph{Limited access to a hull and a taciturn pilot.} \emph{Free Use:} Water crossing scenes begin \textbf{Dominant}. \emph{Surge:} Evade a patrol with a practiced channel; reduce DV by 1 on the escape roll.

  \item[\textbf{Herbal Plot}] \index{herbs}\index{garden} \emph{Drying racks, clay pots, and a steady hand at dawn.} \emph{Free Use:} Remove Fatigue 1 from one character or gain a \emph{Boon} to a single mundane healing attempt this downtime. \emph{Surge:} Produce a rare poultice; convert a Severe healing clock into two Moderate ticks.

  \item[\textbf{Personal Library}] \index{library!personal} \emph{Shelves of annotated tomes and a reading chair.} \emph{Free Use:} Gain a \emph{Boon} on one Lore roll per session. \emph{Surge:} Quote a relevant passage; reduce DV of a related Investigation by 1.

  \item[\textbf{Stable/Coach House}] \index{stable}\index{transport} \emph{Reliable mounts and drivers who ask few questions.} \emph{Free Use:} Land travel scenes begin \textbf{Dominant}. \emph{Surge:} Produce a fresh mount or spare vehicle; ignore the first \emph{exhaustion} complication.

  \item[\textbf{Alchemist's Workshop}] \index{alchemist} \emph{Reagents, retorts, and a suspicious chimney.} \emph{Free Use:} Produce one dose of a common alchemical item between scenes. \emph{Surge:} Brew a specialized compound; gain a \emph{Boon} on the next Craft (Alchemy) roll.

  \item[\textbf{Messenger Birds}] \index{messenger birds}\index{communication} \emph{Trained pigeons or ravens with coded leg bands.} \emph{Free Use:} Send a simple message across moderate distance safely. \emph{Surge:} Receive urgent news now; mitigate 1 SB from \emph{miscommunication}.
\end{description}

\section{Standard Assets (8 XP)}
\index{assets!standard}
Established holdings with regional reach.
\begin{description}
  \item[\textbf{Charter or Minor Title}] \index{charter}\index{title} \emph{Recognized papers and privileges.} \emph{Free Use:} Secure an audience or official hearing; start \textbf{Dominant}. \emph{Surge:} Invoke prerogative to delay immediate sanctions; mitigate 1 SB of civic backlash.

  \item[\textbf{Guild Seat}] \index{guild} \emph{Voting rights and shop‑rights.} \emph{Free Use:} Commission or procurement opens \textbf{Dominant} or DV −1. \emph{Surge:} A guild factor vouches; fill 1 segment on a \emph{Permit/Access} clock.

  \item[\textbf{Market Front}] \index{market}\index{stall} \emph{A fixed stall and regular customers.} \emph{Free Use:} Treat one \emph{Minor} Asset's upkeep as waived this downtime. \emph{Surge:} Liquidate stock; justify on-hand mundane gear within reason.

  \item[\textbf{Informant Ring}] \index{spies}\index{informants} \emph{Eyes and ears on retainer.} \emph{Free Use:} Targeted inquiry begins \textbf{Dominant}. \emph{Surge:} Reveal a hidden schedule or route now; mitigate 1 SB from \emph{ambush/surprise}.

  \item[\textbf{Mercenary Contract (Cap 2)}] \index{mercenaries} \emph{Access to a small trained unit.} \emph{Free Use:} Introduce temporary \emph{off-screen} security that downgrades “raid” to “attempted raid.” \emph{Surge:} One on-screen intervention (one beat) that improves Position for a withdrawal or breach.

  \item[\textbf{Trading Charter}] \index{trade}\index{charter!trading} \emph{Right-of-way at gates and tolls.} \emph{Free Use:} Border/inspection scenes start \textbf{Dominant}. \emph{Surge:} Produce stamped manifests; DV −1 on the next scrutiny roll.

  \item[\textbf{Licensed Library}] \index{library} \emph{Stacks and a learned custodian.} \emph{Free Use:} Deep‑topic research begins \textbf{Dominant} and gains a \emph{Boon}. \emph{Surge:} Cite a precedent to force a pause in debate (one exchange).

  \item[\textbf{Craftsman's Quarter}] \index{workshops}\index{apprentices} \emph{Shop plus apprentices.} \emph{Free Use:} Clear \emph{Compromised} from a crafted item or structure. \emph{Surge:} Rush a bespoke piece; fill 1 segment on a \emph{Build/Repair} clock.

  \item[\textbf{Temple Complex}] \index{temple} \emph{Courts, choirs, charity.} \emph{Free Use:} Among adherents, start \textbf{Dominant} or mitigate 1 SB from \emph{crowd/morale}. \emph{Surge:} Sanctuary: hostile social pressure must escalate openly or wait.

  \item[\textbf{Harbor Office}] \index{harbor}\index{port} \emph{Berths, booms, bells.} \emph{Free Use:} Priority loading/unloading; start \textbf{Dominant}. \emph{Surge:} Hold the chain for one beat; negate a single \emph{pursuit by water} consequence.

  \item[\textbf{Healing House}] \index{healing}\index{infirmary} \emph{Beds, herbs, bells.} \emph{Free Use:} During downtime, clear \emph{Harm 1} or \emph{Fatigue 2} from one ally. \emph{Surge:} Stabilize now; convert a Severe injury consequence into a 4‑segment Recovery clock.

  \item[\textbf{Observatory}] \index{observatory}\index{navigation} \emph{Lenses and ephemerides.} \emph{Free Use:} Predict timing; navigation/timing openers start \textbf{Dominant}. \emph{Surge:} Name a narrow window; DV −1 for an action attempted inside it.

  \item[\textbf{Performance Venue}] \index{theatre}\index{performance} \emph{Stage and regulars.} \emph{Free Use:} Begin Performance with a receptive crowd. \emph{Surge:} Spotlight an ally; shift a crowd one step toward your desired mood.

  \item[\textbf{Hospitality House}] \index{hospitality}\index{guest right} \emph{Beds, bread, rumor.} \emph{Free Use:} Gain a \emph{Protected pause} in parley on your ground. \emph{Surge:} Invoke guest-right; hostile social pressure must step into the open.

  \item[\textbf{Storage Vault}] \index{vault}\index{storage} \emph{Locks and ledgers, damp stone and iron.} \emph{Free Use (per session):} Ignore the first \emph{theft/sabotage} targeting stored goods. \emph{Surge:} Produce a cached item plausibly stored here.

  \item[\textbf{Hidden Cache}] \index{cache}\index{emergency} \emph{A buried lockbox or hollow tree, accessible anywhere in the city.} \emph{Free Use:} Once per session, access emergency gear when cornered; start the scene \textbf{Dominant}. \emph{Surge:} Reveal a larger stash; fill 1 segment on a \emph{Supply} clock.

  \item[\textbf{Signal Relay}] \index{relay}\index{communications} \emph{Fires, flags, drums.} \emph{Free Use:} Coordinated actions across distance start \textbf{Dominant}. \emph{Surge:} Push a ready signal now; mitigate 1 SB from \emph{confusion/lag}.

  \item[\textbf{Merchant Caravan}] \index{caravan}\index{trade} \emph{One wagon, two guards, and a route.} \emph{Free Use:} Trade missions begin \textbf{Dominant}. \emph{Surge:} Call in the caravan early; provide unexpected supplies or reinforcements.

  \item[\textbf{Legal Practice}] \index{law}\index{barrister} \emph{Papers, precedents, and clients.} \emph{Free Use:} Legal proceedings start \textbf{Dominant} or DV −1. \emph{Surge:} Produce a forgotten writ; negate one \emph{legal} complication.

  \item[\textbf{Smuggler's Network}] \index{smugglers}\index{underground} \emph{Contacts and hidden routes.} \emph{Free Use:} Illicit transport begins \textbf{Dominant}. \emph{Surge:} Bypass a checkpoint; mitigate 1 SB from \emph{customs/patrols}.

  \item[\textbf{Crafting Guildhall}] \index{guildhall}\index{crafting} \emph{Tools, masters, and apprentices.} \emph{Free Use:} Complex crafting projects gain a \emph{Boon}. \emph{Surge:} Rush production; fill 1 segment on a \emph{Manufacturing} clock.
\end{description}

\section{Major Assets (12 XP)}
\index{assets!major}
Significant holdings that bend regional outcomes.
\begin{description}
  \item[\textbf{Ward License}] \index{city ward}\index{license} \emph{You run a city district's levers.} \emph{Free Use:} In-ward civic openers start \textbf{Dominant}. \emph{Surge:} Re-route patrols or light—shift a chase route 1 step in your favor.

  \item[\textbf{Exclusive Right (Monopoly)}] \index{monopoly} \emph{Control of a staple (salt, timber, ferries).} \emph{Free Use:} In market scenes, ignore the first \emph{supply} complication. \emph{Surge:} Declare a brief scarcity/surplus to leverage terms; DV −1 on a single negotiation.

  \item[\textbf{Fortress/Stronghold}] \index{fortress}\index{stronghold} \emph{Stone, stores, signals.} \emph{Free Use:} In raid/siege framing, negate the first \emph{breach} consequence. \emph{Surge:} Sound beacons; allies arrive in time to improve Position for withdrawal or stand.

  \item[\textbf{Demesne (Estate)}] \index{estate}\index{demesne} \emph{Fields, mills, tenants.} \emph{Free Use:} Waive minor Supply fallout this arc or treat one additional Minor Asset as having free upkeep. \emph{Surge:} Muster teamsters and hands; fill 1 segment on a \emph{Logistics} clock.

  \item[\textbf{Fleet Share}] \index{fleet} \emph{A cut of hulls and signals.} \emph{Free Use:} Maritime travel starts \textbf{Dominant}. \emph{Surge:} Signal a safe corridor; mitigate 1 SB from \emph{storm/patrol}.

  \item[\textbf{College (University Seat)}] \index{university}\index{college} \emph{Faculty, quad, seal.} \emph{Free Use:} In downtime, reduce \emph{one} skill advancement cost by 1 XP (minimum 1). \emph{Surge:} Convene a colloquy to force a pause in magical proceedings (\S\ref{ch:magic}).

  \item[\textbf{Banking Charter}] \index{bank}\index{letters of credit} \emph{Letters, liens, a counting house.} \emph{Free Use:} Present credit to bypass petty fees and minor delays. \emph{Surge:} Freeze a rival's access for one exchange; mitigate 1 SB from \emph{bribery/extortion}.

  \item[\textbf{Diplomatic Enclave}] \index{diplomacy}\index{enclave} \emph{Flag, immunity, ink.} \emph{Free Use:} In foreign courts, begin \textbf{Dominant}. \emph{Surge:} Invoke safe-conduct; hostile action must escalate openly or withdraw.

  \item[\textbf{Mine/Quarry}] \index{mine}\index{quarry} \emph{Ore, stone, skilled crews.} \emph{Free Use:} Clear \emph{Compromised} on one structure or armament. \emph{Surge:} Deliver timbers/stone now; DV −1 on a \emph{reinforce/repair} roll.

  \item[\textbf{Mining Operation}] \index{mining}\index{resources} \emph{Shafts, carts, and crews.} \emph{Free Use:} Ignore the first \emph{resource shortage} complication this arc. \emph{Surge:} Strike a rich vein; produce unexpected wealth or materials.

  \item[\textbf{Military Outpost}] \index{military}\index{outpost} \emph{Barracks, stores, and command.} \emph{Free Use:} Military operations begin \textbf{Dominant}. \emph{Surge:} Call for reinforcements; improve Position for a tactical withdrawal or assault.

  \item[\textbf{Trading Consortium}] \index{consortium}\index{commerce} \emph{Multiple fronts and shared capital.} \emph{Free Use:} Ignore the first price gouging or supply shortage complication this arc. \emph{Surge:} Coordinate a multi-front deal; DV −1 on a major commercial negotiation.

  \item[\textbf{Religious Order}] \index{religious order}\index{temple} \emph{Chapters, rites, and influence.} \emph{Free Use:} Among believers, start \textbf{Dominant} or mitigate 1 SB from \emph{heresy/doubt}. \emph{Surge:} Call a council; force a pause in hostile religious proceedings.
\end{description}

\section{Specialized Assets}
\index{assets!specialized}
Narrow tools with sharp edges.
\begin{description}
  \item[\textbf{Magical Laboratory} (10 XP)] \index{laboratory!magical}\index{rituals} \emph{Wards, benches, focus crystals.} \emph{Free Use:} Once per session, reduce Ritual Backlash by 1 SB (\S\ref{ch:magic}). \emph{Surge:} Contain an emergent effect for one beat; begin the control roll \textbf{Dominant}.

  \item[\textbf{Training Academy} (12 XP)] \index{academy}\index{training} \emph{Masters and mats.} \emph{Free Use:} In downtime, reduce \emph{one} skill advancement cost by 1 XP (min 1) for a designated trainee. \emph{Surge:} Call in a drilled team to assist off-screen; fill 1 segment on a \emph{Preparation} clock.

  \item[\textbf{Intelligence Bureau} (10 XP)] \index{intelligence}\index{analysis} \emph{Analysts and cut-outs.} \emph{Free Use:} Avoid surprise; the first recon opener starts \textbf{Dominant}. \emph{Surge:} Burn a channel; mitigate 1 SB from \emph{counterintelligence}.
\end{description}

\section{Tier III Institutions (16–20 XP)}
\index{assets!tier III}\index{institutions}
Empire-shaping engines. Each requires ongoing political play.
\begin{description}
  \item[\textbf{Grand Forge-Citadel} (18 XP)] \index{forge}\index{citadel} \emph{City heat and hammer.} \emph{Free Use:} Treat a batch of masterwork gear as immune to the first \emph{gear} complication this arc. \emph{Surge:} Rush a critical piece; fill 2 segments on a \emph{Build/Arm} clock.

  \item[\textbf{Great Library} (16 XP)] \index{great library} \emph{Memory palace of the region.} \emph{Free Use:} Major inquiry begins \textbf{Dominant} and gains a \emph{Boon}. \emph{Surge:} Produce a sealed citation; negate a single \emph{misinterpretation} complication.

  \item[\textbf{Grand Exchange} (20 XP)] \index{exchange}\index{bourse} \emph{Markets move when you breathe.} \emph{Free Use:} In high finance, ignore the first \emph{market shock}. \emph{Surge:} Halt trading bells for one exchange; DV −1 on one decisive deal.

  \item[\textbf{Fleet Command} (18 XP)] \index{admiralty}\index{naval command} \emph{Signals across sea lanes.} \emph{Free Use:} Declare blockade/corridor; hostile egress gains +1 step of difficulty. \emph{Surge:} Redirect a squadron; improve Position for a maritime raid or escape.

  \item[\textbf{Hydraulic Works} (16 XP)] \index{aqueduct}\index{irrigation} \emph{Water is wealth.} \emph{Free Use:} Cancel the first \emph{shortage/drought} fallout this arc. \emph{Surge:} Flood or feed a field at the critical moment; fill 1 segment on a \emph{Famine Relief} clock.

  \item[\textbf{Palace Complex} (20 XP)] \index{palace}\index{court} \emph{Keys to office and omen.} \emph{Free Use:} Political openers start \textbf{Dominant}. \emph{Surge:} Convene a chamber; convert one \emph{scandal} consequence into a public inquiry you can contest.
\end{description}

\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title=Legacy Assets (25+ XP)]
Entire regions, autonomous armies, or a kingdom's treasury. These are campaign anchors, not purchases. Acquire them through story, not XP alone. Examples:
\begin{itemize}
  \item \textbf{March Lordship} – Rule a border territory; must defend it from raids.
  \item \textbf{Fleet of the Inner Sea} – Control over a major naval force; attracts enemies and admirals.
  \item \textbf{Imperial Restoration Charter} – You speak with the voice of a fallen empire, but rivals will never stop scheming.
\end{itemize}
\end{tcolorbox}

\section{Upgrades \& Variants}
\index{assets!upgrades}\index{assets!variants}
Bolster an existing Asset with additional investment (costs stack).
\begin{description}
  \item[\textbf{Enhanced Security} (+2 XP)] \index{security} Hardened doors, sober watchers. \emph{Effect:} Each session, ignore or mitigate 1 SB from theft/sabotage against this Asset.

  \item[\textbf{Expanded Scope} (+3 XP)] \index{scope} Wider footprint or second site. \emph{Effect:} Use the Asset's \emph{Free Use} twice per session \emph{or} across two adjacent districts/routes.

  \item[\textbf{Arcane Wards} (+4 XP)] \index{wards}\index{magic!wards} Sanctified and sigiled. \emph{Effect:} Once per session, reduce magical Backlash by 1 SB on-site \emph{or} treat supernatural intrusion as one step easier to detect.

  \item[\textbf{Political Patrons} (+3 XP)] \index{patrons}\index{politics} Seals and signatures. \emph{Effect:} In civic scenes tied to this Asset, begin \textbf{Dominant} or DV −1.
\end{description}

\section{Creating Your Own Asset}
\index{assets!creation}
When drafting a new Asset, answer:
\begin{enumerate}
  \item \textbf{Scope}: Who benefits? (you / party / named faction)
  \item \textbf{Fictional Edge}: What doors open or dangers dim?
  \item \textbf{Cost Tier}: Minor 4 / Standard 8 / Major 12 / Tier III 16–20.
  \item \textbf{Free Use}: One clear, repeatable off-screen benefit.
  \item \textbf{Scene Surge}: One flashy, plausible \emph{on-screen} move (1 beat) at the cost of 1 Boon.
  \item \textbf{Upkeep Hook}: What happens if you neglect it?
\end{enumerate}

\section{Using Assets in Play}
\index{assets!in play}\index{downtime}\index{flashbacks}
\begin{itemize}
  \item \textbf{Frame the benefit} at scene start (\emph{why is today easier?}).
  \item \textbf{Flash back} with a modest cost to show the Asset's prior work.
  \item \textbf{Tie complications} to Asset condition—if it's \emph{Neglected}, the GM has permission to spend SB against it first.
  \item \textbf{Spotlight spread}: Rotate which Assets take the stage so each investment breathes on-screen now and then.
\end{itemize}

\paragraph{Example of Play}
Rellan uses his \textbf{Harbor Office (Standard Asset)} free use to start a customs inspection at \textbf{Dominant} position. Later, when a rival tries to trap him on the wharf, he spends a Boon for a \textbf{Scene Surge}: the harbor chain is raised, blocking pursuit for one beat—just enough to slip into a waiting skiff.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Asset Quick Reference,fonttitle=\bfseries]
\begin{itemize}
  \item \textbf{Costs}: Minor 4 XP / Standard 8 XP / Major 12 XP / Tier III 16–20 XP.
  \item \textbf{Free Use}: Once per session, off-screen leverage (position, DV −1, fill 1 clock segment, or mitigate 1 SB).
  \item \textbf{Scene Surge}: Spend 1 Boon for a one-beat, on-screen intervention.
  \item \textbf{Upkeep}: Pay XP = base cost \emph{or} devote a downtime scene; neglect degrades condition.
  \item \textbf{Capacity (Optional)}: Assets $\leq$ \textbf{Spirit} are easy; beyond that, upkeep taxes mount.
\end{itemize}
\end{tcolorbox}

\section{Summary}
\index{assets!summary}
Minor Assets grease doors; Standard ones bend routes; Major holdings set tables; Tier III institutions tilt the map. Choose what to own, then let the world answer.

\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title={From Faction to Asset: Taking Over a Terrestrial Patron}]
\index{terrestrial patrons!takeover}\index{assets!from factions}
Sometimes you stop owing the guild – and start owning it. Rising to lead a faction (Thieves’ Guild, Mercenary Company, Banking Charter, etc.) is a major story arc, not a simple XP purchase. This sidebar outlines how to turn a terrestrial patron into a personal Asset.

\paragraph{Narrative Prerequisites}
Before you can attempt a takeover, you must:
\begin{itemize}
  \item Have at least \textbf{6 Obligation} with the faction (you are deeply entangled).
  \item Complete a \textbf{major service} for the faction (GM’s discretion – e.g., saving it from destruction, winning a war, exposing a traitor at great cost).
  \item Secure at least one \textbf{internal ally} with significant influence (another captain, a guildmaster, a wealthy backer).
\end{itemize}

\paragraph{The Takeover Arc}
The GM and player collaborate on a 2–3 session arc that typically involves:
\begin{enumerate}
  \item \textbf{Crisis or Power Vacuum} – The current leader dies, disappears, or is disgraced.
  \item \textbf{Claim of Right} – You present a valid claim (charter, election, duel victory, popular support).
  \item \textbf{Consolidation} – Defeat or co‑opt rivals, bribe or intimidate dissenters, and secure the faction’s resources.
\end{enumerate}

\paragraph{Mechanical Cost}
Once the arc is complete, you must spend \textbf{12 XP} (the cost of a Major Asset) to convert the faction into a personal Asset. This represents the time, bribes, and institutional restructuring required to make it \emph{yours} rather than a temporary alliance.

\paragraph{Converting the Asset}
The new Asset uses the statistics of the most appropriate existing Asset from this compendium, modified by the faction’s nature:
\begin{itemize}
  \item \textbf{Thieves’ Guild} → \textbf{Informant Ring} (Standard) or \textbf{Smuggler’s Network} (Standard)
  \item \textbf{Mercenary Company} → \textbf{Mercenary Contract} (Minor/Standard) or \textbf{Military Outpost} (Major)
  \item \textbf{Banking Charter} → \textbf{Banking Charter} (Major) or \textbf{Trading Consortium} (Major)
  \item \textbf{Royalist Network} → \textbf{Intelligence Bureau} (Specialized) or \textbf{Diplomatic Enclave} (Major)
  \item \textbf{Ecktorian Mystery Cult} → \textbf{Temple Complex} (Standard) or \textbf{Magical Laboratory} (Specialized)
\end{itemize}
The GM may also create a \textbf{custom Asset} using the guidelines in \S\ref{sec:creating-assets}.

\paragraph{What You Gain}
\begin{itemize}
  \item The Asset’s normal \textbf{Free Use} and \textbf{Scene Surge} benefits.
  \item The \textbf{Access Tag} associated with the faction (e.g., \textit{Guild-Ledgered}, \textit{Harbor Handsigns}).
  \item A permanent \textbf{obligation track} toward the faction’s members (now your underlings). This track is smaller (4 segments) and fills when you neglect their welfare or make reckless decisions that endanger them.
\end{itemize}

\paragraph{What You Lose}
\begin{itemize}
  \item You can no longer use the faction as a \emph{terrestrial patron} to gain Obligation‑free perks – you are now responsible for it.
  \item The faction becomes a target for your enemies. The GM may advance a \textbf{Rival Faction clock} whenever you use the Asset aggressively.
  \item You must pay normal \textbf{upkeep} (XP or downtime scenes) to keep the Asset \emph{Maintained}. Neglect causes internal dissent, corruption, or defections.
\end{itemize}

\paragraph{Optional: Gradual Takeover (No XP Cost)}
For a slower, more story‑driven approach, the player can forgo the 12 XP cost but must instead \textbf{spend a downtime scene per session} managing the faction (equivalent to Intensive upkeep) and accept that the Asset’s benefits are unreliable until they have successfully completed a full arc of consolidation. The GM may allow the XP cost to be paid in installments (e.g., 4 XP per downtime period) to represent gradual investment.

\paragraph{Example: Rellan Takes the Black Ledger}
Rellan has served the Black Ledger smuggling syndicate for two arcs, accruing 6 Obligation. After exposing a rival who betrayed the syndicate, the aging leader steps down. Rellan spends 12 XP, completes a short arc consolidating loyal captains, and converts the syndicate into a \textbf{Smuggler’s Network (Standard Asset)}. He now commands hidden routes and safehouses – but must keep the crew paid and protected, or an internal revolt clock will start ticking.
