\chapter{Character Backgrounds}
\label{ch:backgrounds}

% === Epigraph / Intro vignette (place at chapter start or page gutter) ===
\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title={From the journal of a Thepyrgosi Witch Hunter},fonttitle=\bfseries]
They call it mercy when the Tower does it and cruelty when I do. The difference is a seal of wax and a hymn at the gate.

I was a bright spark once—could taste iron on my tongue when the air went thin before a working. The masters in Thepyrgos tested me, weighed me, and found me \emph{nearly} enough. ``Nearly'' is a door that never opens again.

So I learned the other doors.

I keep a ledger: names, places, the hour the lamps guttered in a tenement, the way the brick sweats when a child hums without meaning to. I carry writs that make people look away. I carry cord and a promise I do not speak aloud. The University pays for talent delivered in time; the city pays for fires that never start. Between those accounts, I balance what I can.

You will say I go too far. Sometimes I do. Sometimes the street is a tinderbox and a wrong whisper is a spark. Sometimes a family would rather believe in saints than in ash. I give them a story they can live with and take away the one that might kill them.

When the bells of Thepyrgos answer the river wind, I walk the colonnade to the bursar's door and do not look at the windows. The work is ugly. The work is necessary. The night does not care which of those is true—only that the lamps keep burning.
\end{tcolorbox}

Before your character took their first deliberate risk, they belonged somewhere. Backgrounds are the trail-dust on your boots, the school-ink on your fingers, the bell-pattern you still tap when frightened. They explain how you know what you know—and who still knows \emph{you}.

\section{Understanding Backgrounds}
\index{backgrounds}\index{character creation!backgrounds}

Backgrounds serve several important functions:
\begin{itemize}
\item Provide narrative context for abilities and choices
\item Explain how you gained skills, languages (\S\ref{ch:languages}), and connections
\item Create ties to regions and factions (\S\ref{ch:world-regions})
\item Offer roleplaying hooks, obligations, and privileges
\item Seed reliable \emph{contacts}, \emph{customs}, and \emph{places} that matter on-screen
\end{itemize}

\paragraph{Mechanical Hooks.}
\index{backgrounds!mechanics}
Choose \textbf{one background} at character creation. It grants:
\begin{itemize}
\item \textbf{Access Tags} (narrative permissions; see below)
\item \textbf{A Signature Contact} (named NPC; usually a Cap~1 follower you \emph{do not} pay XP for; they cannot take independent actions, but can grant +1d \emph{once per scene} when their help is plausible)
\item \textbf{A Background Boon} (once per session, gain a +1d or reduce DV by 1 for a task strongly tied to the background's sphere)
\item \textbf{An Obligation Clock} (4 segments; when filled, it creates a \emph{call-in} from that world—debt, favor, duty; see \S\ref{sec:bg-obligation})
\end{itemize}

\paragraph{Access Tags.}
\index{access tags}\index{backgrounds!access}
Tags describe \emph{where} and \emph{with whom} you have standing. Examples: \textit{Guild-Ledgered}, \textit{Court-Announced}, \textit{Veteran-of-the-Marches}, \textit{Bell-Kin of the Causeways}, \textit{Wind-Rider's Tokens}, \textit{Charter-Stamped}. When a scene engages a tag, the GM may:
\begin{itemize}
\item Waive a trivial gate (no roll) or
\item Lower Position risk by one step for first contact or
\item Offer a \emph{Devil's Bargain} tied to that society's expectations
\end{itemize}

\section{Common Background Categories}
\index{backgrounds!categories}

\subsection*{Social Class Backgrounds}
\index{backgrounds!social}\index{aristocratic}\index{merchant}\index{common folk}

\paragraph{Aristocratic House Scion}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Court-Announced}, \textit{Heraldic Right}
\item \textbf{Skills}: Sway, Command, Lore (customs)
\item \textbf{Signature Contact}: Family steward (Cap~1)
\item \textbf{Background Boon}: Once/session in a formal setting, treat Social Position as one step better for an exchange.
\item \textbf{Obligation Seeds}: Marriage alliances, estate audits, house feuds
\item \textbf{Challenges}: Intrigue clocks, reputation stakes
\end{itemize}

\paragraph{Merchant Factor}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Guild-Ledgered}, \textit{Caravan Rights}
\item \textbf{Skills}: Sway (negotiation), Investigation (accounts), Lore (markets)
\item \textbf{Signature Contact}: Bookkeeper or quartermaster (Cap~1)
\item \textbf{Background Boon}: Once/session when buying/selling or logistics, reduce DV by 1 or gain +1d.
\item \textbf{Obligation Seeds}: Notes due, partner demands, embargoes
\end{itemize}

\paragraph{Common Folk}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Neighborhood Trust}, \textit{Guild-Friend}
\item \textbf{Skills}: Craft (one), Endurance, Streetwise or Nature
\item \textbf{Signature Contact}: Elder, foreman, or midwife (Cap~1)
\item \textbf{Background Boon}: Once/session among familiar folk, convert a failure on routine labor to a success-with-cost.
\item \textbf{Obligation Seeds}: Communal favors, festival work, mutual aid
\end{itemize}

\subsection*{Professional Backgrounds}
\index{backgrounds!professional}\index{military}\index{academic}\index{criminal}

\paragraph{Marcher Veteran}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Veteran-of-the-Marches}, \textit{Muster Papers}
\item \textbf{Skills}: Melee or Ranged, Tactics, Endurance
\item \textbf{Signature Contact}: Old sergeant (Cap~1)
\item \textbf{Background Boon}: Once/session in organized violence, ignore the first point of situational Disadvantage from confusion or fear.
\item \textbf{Obligation Seeds}: Recall to colors, feud legacies, back-pay claims
\end{itemize}

\paragraph{Academic of the Dialogues}
\begin{itemize}
\item \textbf{Access Tags}: \textit{School-Sealed}, \textit{Reading Privileges}
\item \textbf{Skills}: Lore, Investigation, Sway (debate)
\item \textbf{Signature Contact}: Lecturer or archivist (Cap~1)
\item \textbf{Background Boon}: Once/session in research or disputation, ask one clarifying question the GM must answer plainly or point to the shelf that does.
\item \textbf{Obligation Seeds}: Citations owed, controversial thesis, patron politics
\end{itemize}

\paragraph{Underworld Hand}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Harbor Handsigns}, \textit{Black Banner Cant}
\item \textbf{Skills}: Stealth, Deception, Streetwise
\item \textbf{Signature Contact}: Fence or fixer (Cap~1)
\item \textbf{Background Boon}: Once/session when using a \emph{cant} (\S\ref{sec:secret-coded}), gain +1d to set up or slip out.
\item \textbf{Obligation Seeds}: Debts, heat with a rival crew, burned safehouse
\end{itemize}

\subsection*{Regional Backgrounds}
\index{backgrounds!regional}\index{urban}\index{rural}\index{frontier}

\paragraph{Urban Child of Silkstrand}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Street-Stall Rights}, \textit{Night-Patrol Familiar}
\item \textbf{Skills}: Streetwise, Perception, Sway (fast talk)
\item \textbf{Signature Contact}: Watch-lantern or stall-keeper (Cap~1)
\item \textbf{Background Boon}: Once/session in a dense city, you know a \emph{shortcut or rumor} that reduces time/cost by one step.
\end{itemize}

\paragraph{Rural Valewood Forager}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Green Courts Token}, \textit{River-Ferry Nod}
\item \textbf{Skills}: Nature, Survival, Lore (spirits)
\item \textbf{Signature Contact}: Leaf-warden (Cap~1)
\item \textbf{Background Boon}: Once/session in the wild, downgrade a Travel complication by one step.
\end{itemize}

\paragraph{Frontier Marcher Homesteader}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Land-Grant Papers}, \textit{Border Moot Voice}
\item \textbf{Skills}: Craft (build), Endurance, Tactics or Diplomacy
\item \textbf{Signature Contact}: Neighboring elder (Cap~1)
\item \textbf{Background Boon}: Once/session at a contested site, shift Position up by one step when defending what's yours.
\end{itemize}

\section{World-Linked Backgrounds}
\index{backgrounds!cultures}\index{cultures!backgrounds}

\paragraph{Aeler Charter Apprentice}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Charter-Stamped}, \textit{Stone-Speech Initiate}
\item \textbf{Skills}: Craft, Lore (engineering), Investigation (plans)
\item \textbf{Signature Contact}: Oathscribe (Cap~1)
\item \textbf{Boon}: Once/session around structures, ignore the first \emph{environmental} complication from collapse or strain.
\end{itemize}

\paragraph{Mistlands Bell-Keeper}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Bell-Law}, \textit{Causeway Rights}
\item \textbf{Skills}: Perception, Lore (omens), Survival (fens)
\item \textbf{Signature Contact}: Dusk-warden (Cap~1)
\item \textbf{Boon}: Once/session, ring a lawful pattern to reframe a fog scene's Position from \emph{Risky} to \emph{Controlled} for one exchange.
\end{itemize}

\paragraph{Ykrul Wind-Runner}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Wind-Knot Tokens}, \textit{Banner Hospitality}
\item \textbf{Skills}: Survival (steppe), Ranged or Riding, Sway (customs)
\item \textbf{Signature Contact}: Banner-aunt (Cap~1)
\item \textbf{Boon}: Once/session in open country, coordinate allies at distance; they gain a one-time \emph{Boon} if they can hear your call.
\end{itemize}

\paragraph{Kahfagian Dock-Priest Acolyte}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Harbor Blessings}, \textit{Beacon Rights}
\item \textbf{Skills}: Lore (sea rites), Sway (crews), Investigation (cargo)
\item \textbf{Signature Contact}: Beacon-keeper (Cap~1)
\item \textbf{Boon}: Once/session at sea or harbor, treat one navigation check as if aided by favorable weather (DV--1).
\end{itemize}

\section{Background Benefits and Limits}
\index{backgrounds!benefits}\index{backgrounds!limitations}

\subsection*{Typical Benefits}
\begin{itemize}
\item \textbf{Skill Access}: Strong narrative justification for certain tests
\item \textbf{Resource Access}: Borrowed tools, rooms, shrines, or records (scene-time instead of XP; see Assets, \S\ref{ch:assets-followers})
\item \textbf{Social Access}: Entry to circles your tag recognizes
\item \textbf{Knowledge}: Clues, maps, or customs that lower DV by 1
\item \textbf{Reputation}: Advantage on \emph{first impressions} in-tag
\end{itemize}

\subsection*{Common Limitations}
\begin{itemize}
\item \textbf{Obligations}: Duties to house, guild, banner, or court
\item \textbf{Restrictions}: Codes, taboos, or public expectations
\item \textbf{Enemies}: Rival houses, crews, or schools
\item \textbf{Secrets}: Leverage others can pull
\item \textbf{Dependencies}: Loss of standing if you neglect ties
\end{itemize}

\section{Custom Backgrounds}
\index{backgrounds!creating}

\subsection*{Build Procedure (Player \& GM)}
Pick \textbf{1 Access Tag}, \textbf{1 Signature Contact}, \textbf{1 Boon}, and \textbf{1 Obligation Seed}. Name places and people.
\begin{itemize}
\item \textbf{Origin}: Where were you taught/raised/forged?
\item \textbf{Training}: Which \emph{two} skills does this background naturally support?
\item \textbf{Experiences}: One memory that still changes how you act
\item \textbf{Relationships}: One ally, one rival; both want something
\item \textbf{Values}: What won't you let slide?
\end{itemize}

\subsection*{Mechanical Considerations}
\index{backgrounds!mechanics mapping}
\begin{itemize}
\item \textbf{Boon Scope}: DV--1 or +1d once/session when strongly on-theme
\item \textbf{Contact Use}: +1d as assist; cannot take independent actions unless purchased as a \emph{Follower} (Cap$^2$ XP; \S\ref{ch:assets-followers})
\item \textbf{Assets}: Borrowing an institutional asset costs \emph{scene time} (counts as upkeep for that asset this downtime) but \(0\)~XP
\item \textbf{Languages}: Choose one tongue linked to the background (see \S\ref{ch:languages}); at least \emph{Conversational}
\end{itemize}

\section{Integration with Development}
\index{backgrounds!integration}\index{advancement!backgrounds}

\subsection*{Early Career}
\begin{itemize}
\item Apprentice, recruit, junior clerk, initiate bell-keeper
\item Boon is modest; obligation triggers are frequent but small
\end{itemize}

\subsection*{Mid-Career}
\begin{itemize}
\item Foreman, sergeant, factor, lecturer
\item Add a \textbf{second Access Tag} or elevate contact to Cap~2 by \emph{paying XP} for a follower
\end{itemize}

\subsection*{Late Career}
\begin{itemize}
\item Master, banner-aunt, beacon-warden, arbiter
\item Background Boon may become a \emph{scene-long permission} at GM approval tied to a vow or festival
\end{itemize}

\section{Background Combinations}
\index{backgrounds!combinations}

\paragraph{Sequential}
\begin{itemize}
\item \textbf{Veteran $\rightarrow$ City Watch}: Access shifts from \textit{Muster Papers} to \textit{Badge Rights}
\item \textbf{Scholar $\rightarrow$ Diplomat}: Keep libraries, gain court
\end{itemize}

\paragraph{Parallel}
\begin{itemize}
\item \textbf{Noble Merchant}: Both \textit{Court-Announced} and \textit{Guild-Ledgered}; obligations multiply
\end{itemize}

\paragraph{Contrasting}
\begin{itemize}
\item \textbf{Criminal Penitent}: Underworld ties vs.\ temple vows; great hooks, sharp bargains
\end{itemize}

\section{Using Backgrounds in Play}
\index{backgrounds!in play}

\subsection*{Player Applications}
\begin{itemize}
\item Name your places: the specific stall, shrine, or yard you know
\item Cash your Boon when you leverage that familiarity
\item Call on your contact to assist or to introduce a new lead
\end{itemize}

\subsection*{GM Applications}
\begin{itemize}
\item Offer choices where background allegiance matters
\item Turn Access Tags into \emph{gates} and \emph{keys}
\item Spend Story Beats (\S\ref{ch:world-interaction}) to \emph{complicate the old ties}
\end{itemize}

\section{Obligations \& Evolution}
\label{sec:bg-obligation}
\index{backgrounds!obligations}\index{backgrounds!evolution}

\subsection*{Obligation Clock (4)}
When filled by misses, bargains, or story, something is \emph{due}.
\begin{description}
\item[1--2] Minor favor, time, or tithe
\item[3] Significant service, costly honesty, dangerous escort
\item[4] Defining ordeal: duel, audit, rite, or trial
\end{description}
Clear segments by honoring duties on-screen, sacrificing resources, or changing affiliation (which creates new clocks).

\subsection*{Enhancement \& Transformation}
\begin{itemize}
\item \textbf{Enhance}: Add a tag or upgrade the contact via XP (Follower rules)
\item \textbf{Transform}: Rewrite the background after a major arc; keep one tag, swap the rest, reset the clock
\item \textbf{Resolution}: Retire a background after paying its \emph{price}; gain a \emph{Legacy Note}—a small permanent permission
\end{itemize}

\section{Tables \& Tools}
\index{backgrounds!tables}\index{random tables}

\subsection*{Quick Background Seeds}
\small
\begin{tabular}{ll}
\toprule
d6 & Seed \\
\midrule
1 & Failed heir with an honest steward and a dishonest cousin \\
2 & Caravan accountant who knows three border tariffs by heart \\
3 & Bell-tender whose dusk pattern saved a village once \\
4 & Charter apprentice who broke a measure and hid it \\
5 & Ykrul courier who owes horses to two banners \\
6 & Dock acolyte who lit the beacon in a storm against orders \\
\bottomrule
\end{tabular}
\normalsize

\subsection*{GM SB Menu: Background Complications}
\index{complications!background}
\begin{itemize}
\item \textbf{1 SB}: A rival claims your tag is invalid here; prove it or pay time
\item \textbf{2 SB}: Your contact is \emph{busy} or \emph{watched}; help will cost you exposure
\item \textbf{3 SB}: Obligation advances; an audit, rite, or muster arrives early
\item \textbf{4 SB}: An enemy weaponizes your background—publicly
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Background Quick Reference,fonttitle=\bfseries]
\textbf{At Creation}
\begin{itemize}
\item Choose: 1 \emph{Access Tag}, 1 \emph{Signature Contact} (Cap~1 assist), 1 \emph{Boon} (once/session), 1 \emph{Obligation Seed}.
\item Pick 1 language tied to the background at \emph{Conversational} or better (\S\ref{ch:languages}).
\end{itemize}

\textbf{In Play}
\begin{itemize}
\item Invoke tags to open doors or shift Position.
\item Call your contact for +1d assist (no independent actions).
\item Track the \emph{Obligation Clock} (4). Pay it on-screen.
\end{itemize}

\textbf{Advancing}
\begin{itemize}
\item Upgrade contact to a \emph{Follower} by paying XP (Cap$^2$).
\item Add a second tag after a major arc or training.
\end{itemize}
\end{tcolorbox}

\section{Worked Examples}
\index{backgrounds!examples}

\subsection*{House Shield (Aristocratic-Military Hybrid)}
\begin{itemize}
\item \textbf{Tags}: \textit{Court-Announced}, \textit{Veteran-of-the-Marches}
\item \textbf{Skills}: Command, Melee, Lore (protocol)
\item \textbf{Contact}: House steward (Cap~1)
\item \textbf{Boon}: Once/session while guarding a principal, treat one consequence as one step less severe.
\item \textbf{Hook}: Duel clock from an old slight; an audit of campaign spoils
\end{itemize}

\subsection*{Ledger Lantern (Merchant-Temple Bridge)}
\begin{itemize}
\item \textbf{Tags}: \textit{Guild-Ledgered}, \textit{Beacon Rights}
\item \textbf{Skills}: Sway, Investigation (accounts), Lore (sea rites)
\item \textbf{Contact}: Beacon-keeper (Cap~1)
\item \textbf{Boon}: Once/session, secure safe harbor or fair rate; DV--1.
\item \textbf{Hook}: Rivals whisper of skimming tithe; prove the books
\end{itemize}

\subsection*{Fog Warden (Mistlands Civic) }
\begin{itemize}
\item \textbf{Tags}: \textit{Bell-Law}, \textit{Causeway Rights}
\item \textbf{Skills}: Perception, Survival, Sway (local law)
\item \textbf{Contact}: Dusk-warden (Cap~1)
\item \textbf{Boon}: Once/session in mist, choose \emph{route-or-safety}: negate a travel delay or downgrade a hazard.
\item \textbf{Hook}: Someone silences the bells; obligation advances
\end{itemize}

\bigskip
\noindent\emph{Let your background do work for you. Speak its language, call its favors, and pay its debts in play.} The world will answer.\end{chapter}