% !TEX root = ../fates_edge_players_guide.tex

\chapter{Followers}
\label{ch:followers}

\begin{multicols}{2}

\index{Assets}\index{Followers}\index{upkeep}\index{obligations}
In \textit{Fate’s Edge}, your reach is measured by more than sword-arms and spell-forms. \textbf{Assets} move the world when you are elsewhere; \textbf{Followers} act at your side when the camera is on you. Both are power—and both are promises you’ll be asked to keep.

\paragraph{A Lantern in the Rain.}
The courier taps twice on the shutters of the riverside inn. Inside, your \emph{Safehouse} (\S\ref{ch:compendium-assets}) has already warmed a room, stocked dry clothes, and left an unmarked letter. Outside, your \emph{Scout} whistles a two-note warning; the watch turns down a side street. Neither scene happens by accident. You paid for it—in XP, in care, and sometimes in blood.

\section{Assets: Off-Screen Influence}
\index{Assets!overview}

\textbf{Assets} are establishments, titles, networks, ships, charters, vaults, shrines, or specialist tools that operate primarily \emph{between scenes}. They alter Position, cancel specific complications, or frame a scene in your favor without needing you on-site.

\subsection*{Acquisition and Costs}
\index{Assets!acquisition}\index{downtime!assets}
Invest XP and time to found or formalize an Asset:
\begin{description}
  \item[Minor (4 XP)] \index{Assets!Minor} 1 day downtime. \emph{Examples:} safehouse, scholar’s cell, barge, petty title, courier route.
  \item[Standard (8 XP)] \index{Assets!Standard} 1 week downtime. \emph{Examples:} guild seat, spy ring, trading charter, workshop block, temple chapter.
  \item[Major (12 XP)] \index{Assets!Major} 1 month downtime. \emph{Examples:} city charter, fortress, university college, monopoly license, regional network.
\end{description}

\subsection*{Condition States}
\index{Assets!condition}
Track an Asset’s operational health:
\begin{description}
  \item[Maintained] Full capability; all text applies.
  \item[Neglected] -1d to \emph{any} use; fiction reflects leaks, dust, or sullen staff.
  \item[Compromised] Unavailable until repaired (downtime scene + test DV 3+, or an appropriate XP spend).
\end{description}

\subsection*{Activation and Use}
\index{Assets!activation}
\begin{itemize}
  \item \textbf{Off-Screen Effect (free, per session):} Each Asset lists a routine benefit you may claim between scenes or as a scene opener.
  \item \textbf{On-Screen Activation (1 Boon):} Pull an Asset into the moment—call in a favor, open the hidden door, flash the seal. The effect must fit the Asset’s scope.
  \item \textbf{Emergency Use (Compromised):} You \emph{may} force a Compromised Asset to act; the GM adds \textbf{+3 CP} to the scene and narrates the messy fallout.
\end{itemize}

\subsection*{Maintenance}
\index{Assets!maintenance}\index{upkeep}
\begin{itemize}
  \item \textbf{Regular Upkeep:} Each downtime, maintain at least one Asset or it slips one step (Maintained \(\rightarrow\) Neglected \(\rightarrow\) Compromised).
  \item \textbf{Recovery:} Neglected \(\rightarrow\) Maintained with a brief scene or 1 XP; Compromised \(\rightarrow\) Neglected with a full downtime action and successful test (DV 3+, skill varies).
  \item \textbf{Portfolio Size:} You can comfortably maintain \(\textbf{Spirit} + 2\) Assets. Each Asset beyond that imposes +1 DV to maintenance tests this downtime. \index{Spirit (attribute)}
\end{itemize}

\paragraph{Regional Color.}
\index{world!regions}\index{Assets!regional}
On the Sun Coast, a \emph{Trading Charter} opens gates at dusk; in the Mistlands, a \emph{Bell-Right} safehouse answers three-tones at night; in the Stone Kingdoms, a \emph{Charter-Script Vault} remembers debts even when stewards die.

\section{Followers: On-Screen Allies}
\index{Followers!overview}

\textbf{Followers} are named allies who stand in the same spotlight you do. They add dice, take risks, carry messages, and complicate your life in the best possible way.

\subsection*{Capability Rating (Cap)}
\index{Followers!Cap}\index{costs!followers}
Followers are rated \textbf{Cap 1–5}. Cost to recruit is \(\textbf{Cap}^2\) XP; 1–3 days of downtime to find, brief, and bind.
\begin{description}
  \item[Cap 1 (4 XP)] Novice or narrow specialist
  \item[Cap 2 (8 XP)] Competent professional
  \item[Cap 3 (12 XP)] Skilled specialist / veteran
  \item[Cap 4 (16 XP)] Expert with reputation
  \item[Cap 5 (20 XP)] Master; scarce and storied
\end{description}
\textbf{Limit:} You can support \(\textbf{Presence} + 1\) Followers without loyalty complications. \index{Presence (attribute)}

\subsection*{Assistance Mechanics}
\index{Followers!assistance}
\begin{itemize}
  \item \textbf{Base Assist:} Add dice equal to \(\min(\text{Cap}, \text{relevant Skill})\).
  \item \textbf{Hard Cap:} Total assist from \emph{all} sources \(\leq\) \textbf{+3d}. \index{dice caps}
  \item \textbf{Specialization:} Some Followers list a \emph{Specialty} granting +1d in a narrow context (still obeys +3d cap).
  \item \textbf{One Pair of Hands:} Only one Follower may assist a given action.
\end{itemize}

\subsection*{Condition: Exposure and Harm}
\index{Followers!Exposure}\index{Followers!Harm}
Track two meters:
\begin{itemize}
  \item \textbf{Exposure}—attention, stress, notoriety, overextension.
  \item \textbf{Harm}—injury, trauma, broken tools, shaken nerve.
\end{itemize}
Each meter uses: 0 None, 1 Light, 2 Moderate (-1d), 3 Severe (-2d \& risk of Compromised), 4 Critical (Out). Followers also have \emph{Maintained/Neglected/Compromised} states mirroring Assets.

\subsection*{Independent Initiative}
\index{Followers!initiative}
Once per scene (party-wide), one Follower can act independently:
\begin{description}
  \item[Scout \& Signal] Shift an ally’s next action to \emph{Controlled}.
  \item[Distract \& Draw] Reduce a Threat/Hunt/Escape clock by 1 tick.
  \item[Fetch \& Carry] Move an object/person through hazard.
  \item[Cover \& Protect] Eat a consequence meant for a PC.
\end{description}
\textbf{Cost:} Mark +1 \emph{Exposure} or \emph{Harm 1} on that Follower.

\subsection*{Upkeep and Risks}
\index{Followers!upkeep}\index{GM!complications (followers)}
\begin{itemize}
  \item \textbf{Maintenance:} Each downtime, pay XP equal to Cap \emph{or} devote a relationship scene. Skip it and they become \emph{Neglected}.
  \item \textbf{Complication Targeting:} When you gain 2+ CP of consequences on an assisted roll, the GM may assign the fallout to the assisting Follower (fiction first).
  \item \textbf{Off-Screen Actions:} Cap 3+ can attempt notable off-screen tasks during downtime; doing so \textbf{generates 1 CP} for the next related scene.
\end{itemize}

\section{Loyalty and Bonds}
\index{Followers!loyalty}\index{Bonds!followers}

\subsection*{Loyalty States}
\begin{description}
  \item[Wary] Costs +1 XP to maintain; needs convincing to risk.
  \item[Steady] Normal; no extra costs or perks.
  \item[Devoted] Once per arc, downgrade a complication that targets them (Major \(\rightarrow\) Minor). \index{arcs!story}
\end{description}

\subsection*{Bond Integration}
\index{Bonds}\index{Boons!bonds}
Tie Followers to your \emph{Bonds}. A Bonded Follower grants +1d when the action speaks to that Bond. Resolving a Follower’s personal quest yields a \emph{Boon} and may bump Loyalty upward.

\section{Advanced Follower Management}
\index{Followers!advancement}\index{Followers!specialization}

\subsection*{Advancement}
On significant shared accomplishments, a Follower may increase Cap by 1 (max PC tier - 1). XP cost equals \(\text{new Cap}^2 - \text{current Cap}^2\). One advancement per major arc.

\subsection*{Common Specializations}
\begin{description}
  \item[Combat Specialist] +1d assisting violent actions; may duel Cap\(\leq\)3 foes.
  \item[Social Specialist] +1d on parley, rumor-work, courtly rites.
  \item[Technical Specialist] +1d on craft/repair; may clear \emph{Compromised} on one item per downtime.
  \item[Magic Specialist] +1d assisting \emph{Weave} (\S\ref{ch:magic}); may absorb 1 SP of backlash once/session. \index{magic!casting loop}
\end{description}

\subsection*{Groups and Cohorts}
\index{Followers!groups}
A squad, crew, or circle can be run as a single \emph{Group Follower} with one Cap, one Exposure track (shared), and distributed Harm. Treat as one assistant in play.

\section{Sample Followers}
\index{Followers!examples}

\subsection*{Cap 1: Street Informant (Wary)}
\textbf{Skills:} Stealth 2, Sway 1, Notice 1 \quad
\textbf{Assist:} +1d to urban intel \\
\textbf{Quirk:} Paid in favors and gossip; hates written notes. \\
\emph{Vignette:} \textit{“Two taps on the shutters if it’s Watch. Three if it’s worse.”}

\subsection*{Cap 3: Veteran Mercenary (Steady)}
\textbf{Skills:} Melee 3, Tactics 2, Athletics 2 \quad
\textbf{Assist:} +3d (capped) on combat/tactics \\
\textbf{Specialty:} Holds lines; can \emph{Cover \& Protect} without taking Harm once/session. \\
\textbf{Quirk:} Won’t fight for “poetry” or “exposure.”

\subsection*{Cap 5: Master Artificer (Devoted)}
\textbf{Skills:} Craft 4, Lore 3, Investigation 3 \quad
\textbf{Assist:} +3d crafting/research (cap rules) \\
\textbf{Off-Screen:} Can produce a minor magical item per downtime (GM frames a hook). \\
\textbf{Quirk:} Collects flawed prototypes “for teaching.”

\section{Asset \& Follower Synergy}
\index{synergy!assets and followers}
\begin{itemize}
  \item \textbf{Workshop + Artificer:} Clear \emph{Compromised} on gear automatically once/session.
  \item \textbf{Spy Ring + Informant:} First recon opener starts \emph{Controlled} \emph{and} yields a concrete lead.
  \item \textbf{Temple Chapter + Social Specialist:} Convert a \emph{Crowd} complication into a \emph{Petition} clock.
\end{itemize}

\section{Integration with Core Systems}
\index{CP economy}\index{Boons}\index{downtime}
\begin{itemize}
  \item \textbf{CP Economy:} Asset use under pressure invites CP spends; Followers can catch those consequences if fictionally exposed.
  \item \textbf{Boons:} Spend to activate an Asset in-scene, rally a Follower from \emph{Moderate} to \emph{Light} condition, or fuel an Independent Initiative.
  \item \textbf{Downtime:} Best place to repair, recruit, advance, and negotiate terms. Tie these scenes to Bonds to harvest story.
\end{itemize}

\section{Tactical and Strategic Play}
\index{strategy!assets}\index{strategy!followers}
\subsection*{Choosing Assets}
\begin{itemize}
  \item \textbf{Complement} weaknesses; don’t double-buy the same tag—take the best one.
  \item \textbf{Map} your routes: ports, passes, bell-precincts; buy where you travel.
  \item \textbf{Plan} a ladder: Minor \(\rightarrow\) Standard \(\rightarrow\) Major that all talk to each other.
\end{itemize}

\subsection*{Building Your Bench}
\begin{itemize}
  \item \textbf{Diversity} beats duplication: one blade, one face, one fixer.
  \item \textbf{Reliability} over raw Cap if the mission needs quiet.
  \item \textbf{Spread risk:} Rotate who takes Exposure to avoid Critical spikes.
\end{itemize}

\section{GM Menus \& Tables}
\index{GM tools}\index{complications!menus}
\subsection*{Quick CP Menu (Assets)}
\begin{itemize}
  \item \textbf{1 CP:} Schedule slips; you arrive late or early (GM’s choice).
  \item \textbf{2 CP:} A gatekeeper wants proof or a costly token.
  \item \textbf{3 CP:} Rival influence undercuts your Asset this scene.
  \item \textbf{4 CP:} Audit, raid, or scandal; Asset flips to \emph{Compromised}.
\end{itemize}

\subsection*{Quick CP Menu (Followers)}
\begin{itemize}
  \item \textbf{1 CP:} Mark \emph{Exposure 1}—they’re noticed by the wrong eyes.
  \item \textbf{2 CP:} \emph{Harm 1} or \emph{Exposure 2}—their nerves fray or they’re bruised.
  \item \textbf{3 CP:} Lose a tool or contact; -1d until replaced.
  \item \textbf{4 CP:} \emph{Compromised}—captured, spooked, or defecting under pressure.
\end{itemize}

\section{Narrative Seeds}
\index{adventure seeds}\index{hooks!assets}\index{hooks!followers}
\paragraph{The Broken Ledger.}
Your \emph{Banking Charter} flags a forged letter-of-mark. Do you protect the account or the client’s life?

\paragraph{The Bell at Dusk.}
A Mistlands \emph{Safehouse} stops answering the three-tone law. The bell rope’s been cut from the outside.

\paragraph{The White Flag.}
Your Cap 3 mercenary refuses to take a fortified bridge without pay in hand. He’s right—someone’s playing both sides.

\section*{Quick Reference}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Assets \& Followers Quick Reference,fonttitle=\bfseries]
\textbf{Assets}
\begin{itemize}
  \item Cost/Time: Minor 4 XP/1 day; Standard 8 XP/1 week; Major 12 XP/1 month.
  \item Use: 1 free off-screen effect/session; 1 Boon to activate in-scene.
  \item Condition: Maintained \(\rightarrow\) Neglected (-1d) \(\rightarrow\) Compromised (repair in downtime).
  \item Portfolio: Maintain up to Spirit + 2 comfortably.
\end{itemize}

\textbf{Followers}
\begin{itemize}
  \item Cost: Cap\(^2\) XP; Limit: Presence + 1 without strain.
  \item Assist: +\(\min(\)Cap, Skill\()\) dice, \textit{max +3d total}.
  \item Independent Action (party 1/scene): Scout \& Signal; Distract \& Draw; Fetch \& Carry; Cover \& Protect (cost: +Exposure or Harm 1).
  \item Tracks: Exposure/Harm 0–4; states mirror Assets.
\end{itemize}
\end{tcolorbox}

\section*{Closing Image}
\index{flourishes}
\textit{A shuttered lamp, a sealed ledger, a scarred gauntlet resting on a map. You don’t move pieces—you invest in people and places. When you step into the street, the city already leans your way.}

\end{multicols}

\end{chapter}
