\chapter{The Legacy Engine}
\label{ch:legacy-engine}

\section*{Succession \& Faction Building}
\addcontentsline{toc}{section}{Succession \& Faction Building}

\begin{quote}
\emph{“Your legacy is not what you build — it is who you build.”}
\end{quote}

The Legacy Engine is a narrative framework that helps your campaign grow across generations without losing emotional continuity. It treats succession as a story event — earned through action, consequence, and choice — rather than a mechanical transaction.

This chapter is for tables where characters retire, die, or pass on their cause to a follower. It is not required for every campaign, but when the moment comes, these tools help the transition feel earned and the new character feel connected.

\section{Core Philosophy}

The Legacy Engine follows three principles:

\begin{enumerate}
  \item \textbf{Narrative Primacy} — Succession happens because the story demands it, not because a clock fills.
  \item \textbf{People Before Power} — A follower becomes a PC because of who they are, not what they can do.
  \item \textbf{Consequences Persist} — Every legacy creates debts, rivals, and unfinished work that the successor must carry.
\end{enumerate}

Legacy is not inherited cleanly. It comes wrapped in obligation.

\section{When Does Succession Happen?}

Succession is not automatic. It emerges from play. Common triggers include:

\begin{itemize}
  \item A PC retires after completing a major arc.
  \item A PC dies or is permanently removed from play.
  \item A player wants to explore a new character but keep the same faction or cause alive.
  \item The campaign shifts to a new generation (years or decades later).
\end{itemize}

In all cases, the transition should be discussed openly at the table. The GM and player collaborate to find the right moment.

\section{Choosing a Successor}

A successor should already exist in the fiction. Typically, this is a \textbf{Follower} (see \S\ref{ch:assets-followers}) with whom the PC has a meaningful relationship. It could also be a family member, a student, a loyal lieutenant, or even a rival who earns respect.

\paragraph{What Makes a Good Successor?}
\begin{itemize}
  \item They have a name, a face, and a voice.
  \item They have acted independently (off‑screen or on‑screen) at least once.
  \item They want something that the PC’s faction or cause can provide.
  \item They have a flaw or burden that will complicate their rise.
\end{itemize}

If no suitable follower exists, the GM and player can create one retroactively — but that character arrives with less established history. The player may need to spend a session or two building them up before succession feels earned.

\section{The Succession Arc}

A smooth transition usually follows three beats. They can span one session or several, depending on the table’s pace.

\subsection{1. The Call}

The follower faces a moment where they must act beyond their role. Examples:
\begin{itemize}
  \item The PC is incapacitated, absent, or compromised.
  \item A decision must be made that no one else can make.
  \item A crisis demands leadership rather than obedience.
\end{itemize}
This scene answers: \textit{Why this person, now?}

\subsection{2. The Trial}

The follower is tested. The trial can be:
\begin{itemize}
  \item A single high‑stakes scene (e.g., a duel, a negotiation, a desperate escape).
  \item A short sequence of escalating challenges (2–3 scenes).
  \item A moral or political decision with lasting impact.
\end{itemize}
Success is not guaranteed. Failure does not end the story — it changes it. A follower who fails may still become a PC, but they carry a scar, a doubt, or a broken promise into their new role.

\subsection{3. The Ascension}

At a dramatically appropriate moment — often during downtime or immediately after a crisis — the follower formally takes up the mantle. This might involve:
\begin{itemize}
  \item A public ceremony (crowning, oath‑swearing, banner‑raising).
  \item A private passing of a symbol (sword, key, ledger).
  \item A quiet acknowledgment (“You’re in charge now.”).
\end{itemize}
The departing PC may be present, or they may be gone. Either way, the scene marks the end of one chapter and the beginning of another.

\section{What the Successor Inherits}

When a follower becomes a PC, they do not simply copy the old character sheet. Instead, they inherit three things:

\subsection{1. Faction Standing}
The successor begins with the same reputation, allies, and enemies that the previous PC cultivated. The GM may adjust specific relationships based on how the succession occurred (e.g., a contested succession might lose one ally but gain a bitter rival).

\subsection{2. One Unresolved Burden}
The player and GM choose one burden that the successor must carry forward. Examples:
\begin{itemize}
  \item \textbf{Oath of Succession}: “You must protect the valley as I did.”
  \item \textbf{Unfinished Task}: “Find my daughter. Tell her I’m sorry.”
  \item \textbf{Rival Claim}: Another faction – or a jealous lieutenant – disputes your right to lead.
\end{itemize}
This burden is a narrative hook, not a mechanical penalty. It gives the new character immediate purpose.

\subsection{3. A Legacy Bond}
The successor has a bond to their predecessor. Once per session, the player may invoke this bond to:
\begin{itemize}
  \item Recall a memory that provides insight.
  \item Receive instinctive guidance in a familiar situation.
  \item Gain a free \textbf{Boon} when acting to uphold the predecessor’s values.
\end{itemize}
This bond fades over time as the successor finds their own identity – but it never vanishes entirely.

\section{Building a Faction Through Succession}

Each time a character passes the torch, the faction (or cause) grows. It accumulates history, traditions, and institutional memory.

\subsection{Growth Through People}
Rather than tracking territory or wealth, track the faction’s \textbf{reach} through the people who are now part of its story. Each successor adds:
\begin{itemize}
  \item A new contact or ally they brought with them.
  \item A new problem they inherited.
  \item A new custom or saying that arose from their trials.
\end{itemize}
The GM keeps a simple list: \textit{“The Banner of the Broken Wheel,” “the annual vigil at Sootfall Abbey,” “the debt to the Mistland ferryman.”} These are the faction’s real assets.

\subsection{Faction Clocks (Optional)}
For tables that want mechanical weight, the faction can track two simple clocks:
\begin{itemize}
  \item \textbf{Growth (6–8)}: When filled, the faction gains a new narrative permission (e.g., a safehouse in a new city, a reliable spy, a permanent shrine).
  \item \textbf{Health (6)}: When filled, a crisis erupts – a rival attacks, a leader dies, a secret is exposed. The current PC must deal with it or watch the faction fracture.
\end{itemize}
These clocks advance only when the fiction demands – not automatically.

\section{Example: The Fallen Captain}

Miranda has played Kael, captain of the Iron Tide mercenary company, for two years. Kael dies defending a village from a wyrm. The player wants to continue the story.

\paragraph{The Successor}
Kael’s second‑in‑command, a grizzled scout named Vesper, has been a follower (Cap 3) for many sessions. Vesper once disobeyed an order to save wounded civilians – a defining moment.

\paragraph{The Call}
At Kael’s funeral pyre, the company looks to Vesper. No one else steps forward.

\paragraph{The Trial}
Vesper must negotiate a new contract with a wary noble who doubts the company’s future without Kael. The player rolls, succeeds with cost – the noble agrees, but demands a hostage.

\paragraph{The Ascension}
Vesper takes up Kael’s banner. The player erases Kael’s character sheet and creates Vesper as a new PC (starting XP based on campaign tier). Vesper inherits:
\begin{itemize}
  \item The Iron Tide as a \textbf{Mercenary Contract (Standard Asset)}.
  \item The burden: \textit{“Find the noble’s missing son – or lose the contract.”}
  \item A legacy bond to Kael: once per session, recall Kael’s advice to gain a Boon when leading the company.
\end{itemize}
The campaign continues without missing a beat.

\section{Campaign End: The Final Legacy}

When the campaign ends, the player may describe one lasting mark the faction leaves on the world. Examples:
\begin{itemize}
  \item A new law named after the first PC.
  \item A shrine where the PC fell, still tended by followers.
  \item A phrase that enters common speech (“Stand like Kael”).
\end{itemize}
The story ends – but the legacy remains.

\begin{quote}
\emph{A character dies. A faction endures. A story echoes forward.}
\end{quote}
