% !TEX root = ../fates_edge_players_guide.tex
\chapter{Compendium of Assets}
\label{ch:compendium-assets}

\index{assets}\index{holdings}\index{influence}
In \textit{Fate’s Edge}, \textbf{Assets} are your off-screen levers—properties, charters, networks, and institutions that change the fiction between scenes. They do not fight beside you (see \S\ref{ch:assets-followers} for \emph{Followers}) but they shape the field you enter.

\section{Design Principles}
\index{assets!design principles}
Assets are written to:
\begin{itemize}
  \item \textbf{Shift fiction} between scenes: open doors, quiet rumors, stock larders, calm crowds.
  \item \textbf{Create hooks}: each Asset suggests opportunities \emph{and} complications if neglected.
  \item \textbf{Stay narrative-first}: prefer position, scale, and access over flat modifiers.
\end{itemize}

Assets do \textbf{not}:
\begin{itemize}
  \item Act inside a scene on their own (Followers do that, \S\ref{ch:assets-followers}). \index{followers|see{assets and followers}}
  \item Print wealth or XP automatically.
  \item Stack identical tags—if two Assets would grant the same edge, take the best one. \index{assets!stacking}
\end{itemize}

\section{Asset Mechanics (Summary)}
\index{assets!mechanics}\index{upkeep}\index{setback points (SB)}
All Assets share the following properties (expanded rules: \S\ref{ch:assets-followers}):
\begin{description}
  \item[Cost] Minor \textbf{4 XP}, Standard \textbf{8 XP}, Major \textbf{12 XP}. \index{assets!minor}\index{assets!standard}\index{assets!major}
  \item[Establishment] 1 day / 1 week / 1 month respectively to become operational.
  \item[Upkeep] Each downtime, either pay XP equal to the Asset’s base cost \emph{or} devote a full scene tending it. Neglect degrades condition (Maintained $\rightarrow$ Neglected $\rightarrow$ Compromised). \index{assets!conditions}\index{assets!maintenance}
  \item[Free Use] Once per session, trigger the listed \emph{off-screen} benefit (framed at scene start or as flashback with GM assent).
  \item[Scene Surge] Spend \textbf{1 Advancement Point} to turn an Asset into a dramatic \emph{on-screen} intervention as described (one beat, fiction permitting). \index{advancement points}
  \item[Extra Use] Spend \textbf{2 XP} to trigger the off-screen benefit an additional time this session.
  \item[Risk Interface] When an Asset covers risk, it typically \textbf{improves Position}, \textbf{lowers DV by 1}, \textbf{fills 1 segment on a relevant clock}, or \textbf{mitigates 1 SB} (Story Beat) from a thematically linked complication. \index{position}\index{difficulty value (DV)}\index{clocks}\index{setback points (SB)}
\end{description}

\paragraph{Capacity (Optional Rule).}\index{assets!capacity}
As a soft cap, a character comfortably maintains a number of Assets up to their \textbf{Spirit} rating. For each Asset beyond this, increase its upkeep by +50\% or require an additional downtime scene to keep it \emph{Maintained}.

\section{Minor Assets (4 XP)}
\index{assets!minor}
Small holdings, charters, or kits that grant reliable leverage.
\begin{description}
  \item[\textbf{Signature Weapon}] \index{signature weapon} A named heirloom or masterpiece. \emph{Free Use:} If gear damage would compromise it, downgrade to a minor flaw and clear with a quiet maintenance scene. \emph{Surge:} Produce it at the perfect moment; improve Position one step for your first attack this scene.
  \item[\textbf{Scholar’s Cell}] \index{scholar’s cell} A quiet room, permits, and a custodian. \emph{Free Use:} Begin one research scene at \textbf{Controlled} or lower DV by 1. \emph{Surge:} Reveal a cited source now; fill 1 segment on a relevant \emph{Information} clock.
  \item[\textbf{Safehouse Network (Quarter)}] \index{safehouse} Two or three bolt-holes in one district. \emph{Free Use:} Start an entry/exit scene \textbf{Controlled}. \emph{Surge:} Produce a hidden egress; convert one \emph{pursuit} consequence into a temporary complication you can interact with.
  \item[\textbf{Artisan’s Bench}] \index{artisan’s bench} Tools for field repairs and fabrication. \emph{Free Use:} Remove \emph{Compromised} from one mundane item between scenes. \emph{Surge:} Improvise a single-use tool; gain a \emph{Boon} on the immediate Craft check.
  \item[\textbf{Courier Chain}] \index{courier network} Runners, dead-drops, and codes in one city. \emph{Free Use:} Move a message or small parcel safely; start the related parley \textbf{Controlled}. \emph{Surge:} A courier appears now; mitigate 1 SB tied to delay.
  \item[\textbf{Small Holding (Farm/Orchard)}] \index{farm}\index{orchard} Steady larder. \emph{Free Use:} Waive a minor Supply complication or clear Fatigue 1 for one PC during downtime. \emph{Surge:} Produce a timely wagon, mule, or crates to justify a plausible cover story.
  \item[\textbf{Neighborhood Inn/Tavern}] \index{inn}\index{tavern} A familiar house of bread and rumor. \emph{Free Use:} First social opener in that neighborhood starts \textbf{Controlled}. \emph{Surge:} Call in the proprietor; shift a hostile crowd one step toward neutral.
  \item[\textbf{Ritual Niche/Shrine}] \index{shrine}\index{ritual niche} A consecrated corner. \emph{Free Use:} Clear Fatigue 1 or mitigate 1 SB from \emph{morale/fear}. \emph{Surge:} Sanctify a doorway; hostile supernatural presence must reveal itself or back away one band (\S\ref{ch:world-interaction}).
  \item[\textbf{River Skiff/Barge Share}] \index{boat}\index{barge} Limited access to a hull and pilot. \emph{Free Use:} Water crossing scenes begin \textbf{Controlled}. \emph{Surge:} Evade a patrol with a practiced channel; lower DV by 1 on the escape roll.
  \item[\textbf{Herbal Plot}] \index{herbs}\index{garden} Beds of restoratives. \emph{Free Use:} Remove Fatigue 1 from one character or gain a \emph{Boon} to a single mundane healing attempt this downtime. \emph{Surge:} Produce a rare poultice; convert a \emph{Severe} healing clock tick into two \emph{Moderate} ticks.
\end{description}

\section{Standard Assets (8 XP)}
\index{assets!standard}
Established holdings with regional reach.
\begin{description}
  \item[\textbf{Charter or Minor Title}] \index{charter}\index{title} Recognized papers and privileges. \emph{Free Use:} Secure an audience or official hearing; start \textbf{Controlled}. \emph{Surge:} Invoke prerogative to delay immediate sanctions; mitigate 1 SB of civic backlash.
  \item[\textbf{Guild Seat}] \index{guild} Voting rights and shop-rights. \emph{Free Use:} Commission or procurement opens \textbf{Controlled} or DV −1. \emph{Surge:} A guild factor vouches; fill 1 segment on a \emph{Permit/Access} clock.
  \item[\textbf{Market Front}] \index{market}\index{stall} A fixed stall and regulars. \emph{Free Use:} Treat one \emph{Minor} Asset’s upkeep as waived this downtime. \emph{Surge:} Liquidate stock; justify on-hand mundane gear within reason.
  \item[\textbf{Informant Ring}] \index{spies}\index{informants} Eyes and ears on retainer. \emph{Free Use:} Targeted inquiry begins \textbf{Controlled}. \emph{Surge:} Reveal a hidden schedule or route now; mitigate 1 SB from \emph{ambush/surprise}.
  \item[\textbf{Mercenary Contract (Cap 2)}] \index{mercenaries} Access to a small trained unit. \emph{Free Use:} Introduce temporary \emph{off-screen} security that downgrades “raid” to “attempted raid.” \emph{Surge:} One on-screen intervention (one beat) that improves Position for a withdrawal or breach.
  \item[\textbf{Trading Charter}] \index{trade}\index{charter!trading} Right-of-way at gates and tolls. \emph{Free Use:} Border/inspection scenes start \textbf{Controlled}. \emph{Surge:} Produce stamped manifests; DV −1 on the next scrutiny roll.
  \item[\textbf{Licensed Library}] \index{library} Stacks and a learned custodian. \emph{Free Use:} Deep-topic research begins \textbf{Controlled} and gains a \emph{Boon}. \emph{Surge:} Cite a precedent to force a pause in debate (one exchange).
  \item[\textbf{Craftsman’s Quarter}] \index{workshops}\index{apprentices} Shop plus apprentices. \emph{Free Use:} Clear \emph{Compromised} from a crafted item or structure. \emph{Surge:} Rush a bespoke piece; fill 1 segment on a \emph{Build/Repair} clock.
  \item[\textbf{Temple Complex}] \index{temple} Courts, choirs, charity. \emph{Free Use:} Among adherents, start \textbf{Controlled} or mitigate 1 SB from \emph{crowd/morale}. \emph{Surge:} Sanctuary: hostile social pressure must escalate openly or wait.
  \item[\textbf{Harbor Office}] \index{harbor}\index{port} Berths, booms, bells. \emph{Free Use:} Priority loading/unloading; start \textbf{Controlled}. \emph{Surge:} Hold the chain for one beat; negate a single \emph{pursuit by water} consequence.
\end{description}

\section{Major Assets (12 XP)}
\index{assets!major}
Significant holdings that bend regional outcomes.
\begin{description}
  \item[\textbf{Ward License}] \index{city ward}\index{license} You run a city district’s levers. \emph{Free Use:} In-ward civic openers start \textbf{Controlled}. \emph{Surge:} Re-route patrols or light—shift a chase route 1 step in your favor.
  \item[\textbf{Exclusive Right (Monopoly)}] \index{monopoly} Control of a staple (salt, timber, ferries). \emph{Free Use:} In market scenes, ignore the first \emph{supply} complication. \emph{Surge:} Declare a brief scarcity/surplus to leverage terms; DV −1 on a single negotiation.
  \item[\textbf{Fortress/Stronghold}] \index{fortress}\index{stronghold} Stone, stores, signals. \emph{Free Use:} In raid/siege framing, negate the first \emph{breach} consequence. \emph{Surge:} Sound beacons; allies arrive in time to improve Position for withdrawal or stand.
  \item[\textbf{Demesne (Estate)}] \index{estate}\index{demesne} Fields, mills, tenants. \emph{Free Use:} Waive minor Supply fallout this arc or treat one additional Minor Asset as having free upkeep. \emph{Surge:} Muster teamsters and hands; fill 1 segment on a \emph{Logistics} clock.
  \item[\textbf{Fleet Share}] \index{fleet} A cut of hulls and signals. \emph{Free Use:} Maritime travel starts \textbf{Controlled}. \emph{Surge:} Signal a safe corridor; mitigate 1 SB from \emph{storm/patrol}.
  \item[\textbf{College (University Seat)}] \index{university}\index{college} Faculty, quad, seal. \emph{Free Use:} In downtime, reduce \emph{one} Skill advancement cost by 1 (min 1). \emph{Surge:} Convene a colloquy to force a pause in magical proceedings (\S\ref{ch:magic}).
  \item[\textbf{Banking Charter}] \index{bank}\index{letters of credit} Letters and liens. \emph{Free Use:} Present credit to bypass petty fees and minor delays. \emph{Surge:} Freeze a rival’s access for one exchange; mitigate 1 SB from \emph{bribery/extortion}.
  \item[\textbf{Diplomatic Enclave}] \index{diplomacy}\index{enclave} Flag, immunity, ink. \emph{Free Use:} In foreign courts, begin \textbf{Controlled}. \emph{Surge:} Invoke safe-conduct; hostile action must escalate openly or withdraw.
  \item[\textbf{Mine/Quarry}] \index{mine}\index{quarry} Ore, stone, skilled crews. \emph{Free Use:} Clear \emph{Compromised} on one structure or armament. \emph{Surge:} Deliver timbers/stone now; DV −1 on a \emph{reinforce/repair} roll.
\end{description}

\section{Tier III Institutions (16–20 XP)}
\index{assets!tier III}\index{institutions}
Empire-shaping engines. Each requires ongoing political play.
\begin{description}
  \item[\textbf{Grand Forge-Citadel} (18 XP)] \index{forge}\index{citadel} City heat and hammer. \emph{Free Use:} Treat a batch of masterwork gear as immune to the first \emph{gear} complication this arc. \emph{Surge:} Rush a critical piece; fill 2 segments on a \emph{Build/Arm} clock.
  \item[\textbf{Great Library} (16 XP)] \index{great library} Memory palace of the region. \emph{Free Use:} Major inquiry begins \textbf{Controlled} and gains a \emph{Boon}. \emph{Surge:} Produce a sealed citation; negate a single \emph{misinterpretation} complication.
  \item[\textbf{Grand Exchange} (20 XP)] \index{exchange}\index{bourse} Markets move when you breathe. \emph{Free Use:} In high finance, ignore the first \emph{market shock}. \emph{Surge:} Halt trading bells for one exchange; DV −1 on one decisive deal.
  \item[\textbf{Fleet Command} (18 XP)] \index{admiralty}\index{naval command} Signals across sea lanes. \emph{Free Use:} Declare blockade/corridor; hostile egress gains +1 step of difficulty. \emph{Surge:} Redirect a squadron; improve Position for a maritime raid or escape.
  \item[\textbf{Hydraulic Works} (16 XP)] \index{aqueduct}\index{irrigation} Water is wealth. \emph{Free Use:} Cancel the first \emph{shortage/drought} fallout this arc. \emph{Surge:} Flood or feed a field at the critical moment; fill 1 segment on a \emph{Famine Relief} clock.
  \item[\textbf{Palace Complex} (20 XP)] \index{palace}\index{court} Keys to office and omen. \emph{Free Use:} Political openers start \textbf{Controlled}. \emph{Surge:} Convene a chamber; convert one \emph{scandal} consequence into a public inquiry you can contest.
\end{description}

\section{Specialized Assets}
\index{assets!specialized}
Narrow tools with sharp edges.
\begin{description}
  \item[\textbf{Magical Laboratory} (10 XP)] \index{laboratory!magical}\index{rituals} Wards, benches, focus. \emph{Free Use:} Once per session, reduce Ritual Backlash by 1 SB (\S\ref{ch:magic}). \emph{Surge:} Contain an emergent effect for one beat; begin the control roll \textbf{Controlled}.
  \item[\textbf{Healing House} (8 XP)] \index{healing}\index{infirmary} Beds, herbs, bells. \emph{Free Use:} During downtime, clear \emph{Harm 1} or \emph{Fatigue 2} from one ally. \emph{Surge:} Stabilize now; convert a \emph{Severe} consequence into a clock you can work.
  \item[\textbf{Observatory} (8 XP)] \index{observatory}\index{navigation} Lenses and ephemerides. \emph{Free Use:} Predict timing; navigation/timing openers start \textbf{Controlled}. \emph{Surge:} Name a narrow window; DV −1 for an action attempted inside it.
  \item[\textbf{Performance Venue} (6 XP)] \index{theatre}\index{performance} Stage and regulars. \emph{Free Use:} Begin Performance with a receptive crowd. \emph{Surge:} Spotlight an ally; shift a crowd one step toward your desired mood.
  \item[\textbf{Intelligence Bureau} (10 XP)] \index{intelligence}\index{analysis} Analysts and cut-outs. \emph{Free Use:} Avoid surprise; the first recon opener starts \textbf{Controlled}. \emph{Surge:} Burn a channel; mitigate 1 SB from \emph{counterintelligence}.
  \item[\textbf{Hospitality House} (6 XP)] \index{hospitality}\index{guest right} Beds, bread, rumor. \emph{Free Use:} Gain a \emph{Protected pause} in parley on your ground. \emph{Surge:} Invoke guest-right; hostile social pressure must step into the open.
  \item[\textbf{Storage Vault} (6 XP)] \index{vault}\index{storage} Locks and ledgers. \emph{Free Use:} Ignore the first \emph{theft/sabotage} targeting stored goods. \emph{Surge:} Produce a cached item plausibly stored here.
  \item[\textbf{Signal Relay} (8 XP)] \index{relay}\index{communications} Fires, flags, drums. \emph{Free Use:} Coordinated actions across distance start \textbf{Controlled}. \emph{Surge:} Push a ready signal now; mitigate 1 SB from \emph{confusion/lag}.
  \item[\textbf{Training Academy} (12 XP)] \index{academy}\index{training} Masters and mats. \emph{Free Use:} In downtime, reduce \emph{one} Skill advancement cost by 1 (min 1) for a designated trainee. \emph{Surge:} Call in a drilled team to assist off-screen; fill 1 segment on a \emph{Preparation} clock.
\end{description}

\section{Upgrades \& Variants}
\index{assets!upgrades}\index{assets!variants}
Bolster an existing Asset with additional investment (costs stack).
\begin{description}
  \item[\textbf{Enhanced Security} (+2 XP)] \index{security} Hardened doors, sober watchers. \emph{Effect:} Each session, ignore or mitigate 1 SB from theft/sabotage against this Asset.
  \item[\textbf{Expanded Scope} (+3 XP)] \index{scope} Wider footprint or second site. \emph{Effect:} Use the Asset’s \emph{Free Use} twice per session \emph{or} across two adjacent districts/routes.
  \item[\textbf{Arcane Wards} (+4 XP)] \index{wards}\index{magic!wards} Sanctified and sigiled. \emph{Effect:} Once per session, reduce magical Backlash by 1 SB on-site \emph{or} treat supernatural intrusion as one step easier to detect.
  \item[\textbf{Political Patrons} (+3 XP)] \index{patrons}\index{politics} Seals and signatures. \emph{Effect:} In civic scenes tied to this Asset, begin \textbf{Controlled} or DV −1.
\end{description}

\section{Creating Your Own Asset}
\index{assets!creation}
When drafting a new Asset, answer:
\begin{enumerate}
  \item \textbf{Scope}: Who benefits? (you / party / named faction)
  \item \textbf{Fictional Edge}: What doors open or dangers dim?
  \item \textbf{Cost Tier}: Minor 4 / Standard 8 / Major 12 / Tier III 16–20.
  \item \textbf{Free Use}: One clear, repeatable off-screen benefit.
  \item \textbf{Scene Surge}: One flashy, plausible \emph{on-screen} move (1 beat) at the cost of 1 Advancement Point.
  \item \textbf{Upkeep Hook}: What happens if you neglect it?
\end{enumerate}

\section{Using Assets in Play}
\index{assets!in play}\index{downtime}\index{flashbacks}
\begin{itemize}
  \item \textbf{Frame the benefit} at scene start (\emph{why is today easier?}).
  \item \textbf{Flash back} with a modest cost to show the Asset’s prior work.
  \item \textbf{Tie complications} to Asset condition—if it’s \emph{Neglected}, the GM has permission to spend SB against it first.
  \item \textbf{Spotlight spread}: Rotate which Assets take the stage so each investment breathes on-screen now and then.
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Asset Quick Reference,fonttitle=\bfseries]
\begin{itemize}
  \item \textbf{Costs}: Minor 4 XP / Standard 8 XP / Major 12 XP / Tier III 16–20 XP.
  \item \textbf{Free Use}: Once per session, off-screen leverage (position, DV −1, fill 1 clock segment, or mitigate 1 SB).
  \item \textbf{Scene Surge}: Spend 1 Advancement Point for a one-beat, on-screen intervention.
  \item \textbf{Upkeep}: Pay XP = base cost \emph{or} devote a downtime scene; neglect degrades condition.
  \item \textbf{Capacity (Optional)}: Assets $\leq$ \textbf{Spirit} are easy; beyond that, upkeep taxes mount.
\end{itemize}
\end{tcolorbox}

\section{Summary}
\index{assets!summary}
Minor Assets grease doors; Standard ones bend routes; Major holdings set tables; Tier III institutions tilt the map. Choose what to own, then let the world answer.
