\chapter{Followers}
\label{ch:followers}

\index{Assets}\index{Followers}\index{upkeep}\index{obligations}
In \textit{Fate's Edge}, your reach is measured by more than sword-arms and spell-forms. \textbf{Assets} move the world when you are elsewhere; \textbf{Followers} act at your side when the camera is on you. Both are power—and both are promises you'll be asked to keep.

\paragraph{A Lantern in the Rain.}
The courier taps twice on the shutters of the riverside inn. Inside, your \emph{Safehouse} (\S\ref{ch:compendium-assets}) has already warmed a room, stocked dry clothes, and left an unmarked letter. Outside, your \emph{Scout} whistles a two-note warning; the watch turns down a side street. Neither scene happens by accident. You paid for it—in XP, in care, and sometimes in blood.

\section{Assets: Off-Screen Influence}
\index{Assets!overview}

\textbf{Assets} are establishments, titles, networks, ships, charters, vaults, shrines, or specialist tools that operate primarily \emph{between scenes}. They alter Position, cancel specific complications, or frame a scene in your favor without needing you on-site.

\subsection*{Acquisition and Costs}
\index{Assets!acquisition}\index{downtime!assets}
Invest XP and time to found or formalize an Asset:
\begin{description}
  \item[Minor (4 XP)] \index{Assets!Minor} 1 day downtime. \emph{Examples:} safehouse, scholar's cell, barge, petty title, courier route.
  \item[Standard (8 XP)] \index{Assets!Standard} 1 week downtime. \emph{Examples:} guild seat, spy ring, trading charter, workshop block, temple chapter.
  \item[Major (12 XP)] \index{Assets!Major} 1 month downtime. \emph{Examples:} city charter, fortress, university college, monopoly license, regional network.
\end{description}

\subsection*{Condition States}
\index{Assets!condition}
Track an Asset's operational health:
\begin{description}
  \item[Maintained] Full capability; all text applies.
  \item[Neglected] –1d to \emph{any} use; fiction reflects leaks, dust, or sullen staff.
  \item[Compromised] Unavailable until repaired (downtime scene + test DV 3+, or an appropriate XP spend).
\end{description}

\subsection*{Activation and Use}
\index{Assets!activation}
\begin{itemize}
  \item \textbf{Off-Screen Effect (free, per session):} Each Asset lists a routine benefit you may claim between scenes or as a scene opener.
  \item \textbf{On-Screen Activation (1 Boon):} Pull an Asset into the moment—call in a favor, open the hidden door, flash the seal. The effect must fit the Asset's scope.
  \item \textbf{Emergency Use (Compromised):} You \emph{may} force a Compromised Asset to act; the GM adds \textbf{+3 SB} to the scene and narrates the messy fallout.
\end{itemize}

\subsection*{Maintenance}
\index{Assets!maintenance}\index{upkeep}
\begin{itemize}
  \item \textbf{Regular Upkeep:} Each downtime, maintain at least one Asset or it slips one step (Maintained \(\rightarrow\) Neglected \(\rightarrow\) Compromised).
  \item \textbf{Recovery:} Neglected \(\rightarrow\) Maintained with a brief scene or 1 XP; Compromised \(\rightarrow\) Neglected with a full downtime action and successful test (DV 3+, skill varies).
  \item \textbf{Portfolio Size:} You can comfortably maintain a number of Assets equal to your \textbf{Spirit} rating. Each Asset beyond that imposes +1 DV to maintenance tests this downtime. (The GM may increase this limit for high‑reputation campaigns.) \index{Spirit (attribute)}
\end{itemize}

\paragraph{Regional Color.}
\index{world!regions}\index{Assets!regional}
On the Sun Coast, a \emph{Trading Charter} opens gates at dusk; in the Mistlands, a \emph{Bell-Right} safehouse answers three-tones at night; in the Stone Kingdoms, a \emph{Charter-Script Vault} remembers debts even when stewards die.

\section{Followers: On-Screen Allies}
\index{Followers!overview}

\textbf{Followers} are named allies who stand in the same spotlight you do. They add dice, take risks, carry messages, and complicate your life in the best possible way.

\subsection*{Capability Rating (Cap)}
\index{Followers!Cap}\index{costs!followers}
Followers are rated \textbf{Cap 1–5}. Cost to recruit is \(\textbf{Cap}^2\) XP; 1–3 days of downtime to find, brief, and bind.
\begin{description}
  \item[Cap 1 (1 XP)] Novice or narrow specialist
  \item[Cap 2 (4 XP)] Competent professional
  \item[Cap 3 (9 XP)] Skilled specialist / veteran
  \item[Cap 4 (16 XP)] Expert with reputation
  \item[Cap 5 (25 XP)] Master; scarce and storied
\end{description}
\textbf{Limit:} You can comfortably maintain up to \(\textbf{Presence} + 1\) Followers without loyalty complications. Exceeding this limit increases the chance of conflict or neglect (GM’s discretion). \index{Presence (attribute)}

\subsection*{Recruiting Followers}
\index{Followers!recruitment}
Recruitment is a narrative scene, not a menu purchase. The GM should frame a short scene where you encounter a potential follower, negotiate terms, and prove your worth. Common hooks:
\begin{itemize}
  \item You save a life; they offer service to repay a debt.
  \item You share a common enemy; they join to settle a grudge.
  * They are down on their luck; you offer purpose and pay.
  \item They are an old contact from your background; you call in a favor.
\end{itemize}
Once terms are agreed, you pay XP and downtime as above. The GM may waive the XP cost if the follower’s motivation is purely narrative (e.g., a rescued prisoner who becomes a loyal ally). The new follower enters play with \emph{Steady} loyalty and no Exposure or Harm.

\subsection*{Assistance Mechanics}
\index{Followers!assistance}
\begin{itemize}
  \item \textbf{Base Assist:} Add dice equal to \(\min(\text{Cap}, \text{relevant Skill})\) to the PC’s roll.
  \item \textbf{Hard Cap:} Total assist dice from \emph{all} sources (other PCs, followers, assets) \(\leq\) \textbf{+3d}. \index{dice caps}
  \item \textbf{Specialization:} Some Followers list a \emph{Specialty} granting +1d in a narrow context (still obeys +3d cap).
  \item \textbf{One Pair of Hands:} Only one Follower may assist a given action.
\end{itemize}

\subsection*{Condition: Exposure and Harm}
\index{Followers!Exposure}\index{Followers!Harm}
Track two meters, each on a 0–4 scale (0=None, 1=Light, 2=Moderate, 3=Severe, 4=Critical). Each level of Moderate or Severe imposes a –1d penalty to the follower’s own actions (but not to assist dice). Critical means the follower is out of action (captured, unconscious, fled).
\begin{itemize}
  \item \textbf{Exposure} — attention, stress, notoriety, overextension.
  \item \textbf{Harm} — injury, trauma, broken tools, shaken nerve.
\end{itemize}
Followers also have the same \emph{Maintained/Neglected/Compromised} states as Assets, representing their availability and morale.

\subsection*{Independent Initiative}
\index{Followers!initiative}
Once per scene (party-wide), one Follower can act independently instead of assisting. Choose one:
\begin{description}
  \item[Scout \& Signal] Shift an ally’s next action to \emph{Dominant} position.
  \item[Distract \& Draw] Reduce a Threat/Hunt/Escape clock by 1 tick.
  \item[Fetch \& Carry] Move an object or person through a hazard safely.
  \item[Cover \& Protect] Eat a consequence meant for a PC; the follower takes Harm 1 or Exposure 2 (GM’s choice).
\end{description}
\textbf{Cost:} Mark +1 \emph{Exposure} or \emph{Harm 1} on that Follower as described.

\subsection*{Upkeep and Risks}
\index{Followers!upkeep}\index{GM!complications (followers)}
\begin{itemize}
  \item \textbf{Maintenance:} Each downtime, pay XP equal to Cap \emph{or} devote a relationship scene. Skip it and they become \emph{Neglected} (lose 1 loyalty step). Two consecutive neglected downturns may cause them to leave.
  \item \textbf{Complication Targeting:} When you gain 2+ SB of consequences on an assisted roll, the GM may assign the fallout to the assisting Follower (fiction first).
  \item \textbf{Off-Screen Actions:} A Cap 3+ follower can attempt a notable off-screen task during downtime; doing so \textbf{generates 1 SB} for the next related scene (the GM determines the nature of the complication).
\end{itemize}

\section{Loyalty and Bonds}
\index{Followers!loyalty}\index{Bonds!followers}

\subsection*{Loyalty States}
\begin{description}
  \item[Wary] Requires +1 XP upkeep; may refuse dangerous orders.
  \item[Steady] Normal; no extra costs or perks.
  \item[Devoted] Once per arc, downgrade a complication that targets them (Major \(\rightarrow\) Minor). \index{arcs!story}
\end{description}
Loyalty shifts based on treatment, shared victories, and personal quests. A follower who is consistently paid, respected, and protected will rise (Wary → Steady after 2–3 sessions; Steady → Devoted after completing a personal quest). One who is ignored, underpaid, or sacrificed will fall.

\subsection*{Bond Integration}
\index{Bonds}\index{Boons!bonds}
If a Follower is also a \textbf{Bond} (see \S\ref{ch:backgrounds}), they grant the bonded Boon (once per session, +1 Boon when you act to uphold the bond) in addition to their assist dice. Working together to resolve the follower’s personal quest counts as upholding the bond.

\subsection*{Follower Personal Quests}
\index{Followers!personal quests}
Every follower has a hidden (or open) goal. Discover it through play. When you help them achieve it, they:
\begin{itemize}
  \item Gain a permanent loyalty boost (Steady → Devoted).
  \item May unlock a new specialization (GM’s discretion).
  \item You earn a \textbf{Legacy Point} (see \S\ref{ch:legacy-engine} for use).
\end{itemize}
Examples:
\begin{itemize}
  \item Avenge a murdered family member.
  \item Recover a lost heirloom or debt note.
  \item Prove themselves worthy of a guild title.
  \item Find a cure for a chronic illness.
  \item Escape a past that still hunts them.
\end{itemize}

\section{Advanced Follower Management}
\index{Followers!advancement}\index{Followers!specialization}

\subsection*{Advancement}
On significant shared accomplishments, a Follower may increase Cap by 1 (max PC tier – 1). XP cost equals \(\text{new Cap}^2 - \text{current Cap}^2\). One advancement per major arc.

\subsection*{Common Specializations}
\begin{description}
  \item[Combat Specialist] +1d assisting violent actions; may duel Cap\(\leq\)3 foes.
  \item[Social Specialist] +1d on parley, rumor-work, courtly rites.
  \item[Technical Specialist] +1d on craft/repair; may clear \emph{Compromised} on one item per downtime.
  \item[Magic Specialist] +1d assisting \emph{Weave} (\S\ref{ch:magic}); may absorb 1 SB of backlash once/session. \index{magic!casting loop}
\end{description}

\subsection*{Groups and Cohorts}
\index{Followers!groups}
A squad, crew, or circle can be run as a single \emph{Group Follower} with one Cap, one Exposure track (shared), and distributed Harm. Treat as one assistant in play. When the group takes Harm, it represents casualties or loss of cohesion; the group may be reduced in Cap temporarily or permanently.

\subsection*{Retirement and Legacy}
\index{Followers!retirement}
A follower may retire from active duty after completing their personal quest or suffering critical harm. They become an \textbf{Off-Screen Asset} (Minor for Cap 1–2, Standard for Cap 3+, free of XP cost) representing their ongoing network or expertise. They may also become a new PC via the Legacy Engine (\S\ref{ch:legacy-engine}).

\section{Sample Followers}
\index{Followers!examples}

\subsection*{Cap 1: Street Informant (Wary)}
\textbf{Skills:} Stealth 2, Sway 1, Notice 1 \quad
\textbf{Assist:} +1d to urban intel \\
\textbf{Specialty:} Knows back ways; reduces escape clock DV by 1 once/session. \\
\textbf{Quirk:} Paid in favors and gossip; hates written notes. \\
\emph{Vignette:} \textit{``Two taps on the shutters if it's Watch. Three if it's worse.''}

\subsection*{Cap 2: Trusted Bodyguard (Steady)}
\textbf{Skills:} Melee 2, Athletics 2, Notice 2 \quad
\textbf{Assist:} +2d on protection/combat \\
\textbf{Specialty:} Can \emph{Cover \& Protect} without taking Harm once/session. \\
\textbf{Quirk:} Never speaks first; always positions themselves between you and exits.

\subsection*{Cap 3: Veteran Mercenary (Steady)}
\textbf{Skills:} Melee 3, Tactics 2, Athletics 2 \quad
\textbf{Assist:} +3d (capped) on combat/tactics \\
\textbf{Specialty:} Holds lines; can \emph{Cover \& Protect} without taking Harm once/session. \\
\textbf{Quirk:} Won't fight for "poetry" or "exposure."

\subsection*{Cap 4: Guild Factor (Devoted)}
\textbf{Skills:} Sway 3, Investigation 3, Lore 2 \quad
\textbf{Assist:} +3d on commerce/legal matters \\
\textbf{Off-Screen:} Can resolve one minor bureaucratic complication per downtime. \\
\textbf{Quirk:} Keeps meticulous records; remembers every favor and debt.

\subsection*{Cap 5: Master Artificer (Devoted)}
\textbf{Skills:} Craft 4, Lore 3, Investigation 3 \quad
\textbf{Assist:} +3d crafting/research (cap rules) \\
\textbf{Off-Screen:} Can produce a minor magical item per downtime (GM frames a hook). \\
\textbf{Quirk:} Collects flawed prototypes "for teaching."

\section{Strategic Play with Followers}

\subsection*{Choosing Followers}
\begin{itemize}
  \item \textbf{Diversity beats duplication:} one combat, one social, one technical.
  \item \textbf{Reliability over raw Cap:} A steady Cap 2 is worth more than a wary Cap 4.
  \item \textbf{Spread risk:} Rotate who takes Exposure to avoid Critical spikes.
  \item \textbf{Bond synergy:} Recruit followers who share your character’s Bonds for bonus Boon generation.
\end{itemize}

\subsection*{Keeping Followers Alive}
\begin{itemize}
  \item \textbf{Use Cover \& Protect sparingly:} It keeps you alive but burns followers.
  \item \textbf{Clear Exposure early:} A single downtime scene can reduce Exposure from 2 to 0.
  \item \textbf{Invest in healing:} A follower with Harm 2 is a liability. Pay for treatment or use a magic specialist.
  \item \textbf{Know when to retreat:} Losing a follower is worse than losing a battle.
\end{itemize}

\subsection*{Leveraging Independent Initiative}
\begin{itemize}
  \item \textbf{Scout \& Signal is free Position:} Use it before a big fight or chase.
  \item \textbf{Distract \& Draw on clocks:} Reduce a pursuit or alarm clock by 1 – often worth more than a +1d.
  \item \textbf{Fetch \& Carry can solve puzzles:} Move a key item across a chasm while you fight.
  \item \textbf{Cover \& Protect as last resort:} Better than you dying, but expensive.
\end{itemize}

\section{Asset \& Follower Synergy}
\index{synergy!assets and followers}
\begin{itemize}
  \item \textbf{Workshop + Artificer:} Clear \emph{Compromised} on gear automatically once/session.
  \item \textbf{Spy Ring + Informant:} First recon opener starts \emph{Dominant} \emph{and} yields a concrete lead.
  \item \textbf{Temple Chapter + Social Specialist:} Convert a \emph{Crowd} complication into a \emph{Petition} clock.
  \item \textbf{Healing House + Combat Specialist:} Recover 1 Harm on a follower between sessions.
\end{itemize}

\section{Integration with Core Systems}
\index{SB economy}\index{Boons}\index{downtime}
\begin{itemize}
  \item \textbf{SB Economy:} Asset use under pressure invites SB spends; Followers can catch those consequences if fictionally exposed.
  \item \textbf{Boons:} Spend to activate an Asset in-scene, rally a Follower from \emph{Moderate} to \emph{Light} condition, or fuel an Independent Initiative.
  \item \textbf{Downtime:} Best place to repair, recruit, advance, and negotiate terms. Tie these scenes to Bonds to harvest story.
\end{itemize}

\section{GM Tools for Followers}

\subsection*{Quick SB Menu (Followers)}
\begin{itemize}
  \item \textbf{1 SB:} Mark \emph{Exposure 1}—they're noticed by the wrong eyes.
  \item \textbf{2 SB:} \emph{Harm 1} or \emph{Exposure 2}—their nerves fray or they're bruised.
  \item \textbf{3 SB:} Lose a tool or contact; –1d until replaced.
  \item \textbf{4 SB:} \emph{Compromised}—captured, spooked, or defecting under pressure (but not killed unless the table has established that as a risk).
\end{itemize}

\subsection*{Follower Personal Quest Seeds}
\index{adventure seeds!follower quests}
\begin{itemize}
  \item \textbf{The Lost Ledger:} Your factor’s father disappeared with a book of debts. Find it before his rivals do.
  \item \textbf{The Broken Oath:} Your mercenary deserted a previous contract; now a bounty hunter seeks them. Protect them or negotiate a pardon.
  \item \textbf{The Silent Bell:} Your scout’s family village stopped answering the three-tone law. Something in the mist took them.
  \item \textbf{The Artificer’s Rival:} The master artificer’s old apprentice is selling flawed enchantments under their name. Clear the reputation.
\end{itemize}

\subsection*{Challenging Followers Without Killing Them}
\begin{itemize}
  \item \textbf{Capture instead of kill:} Rivals want leverage, not corpses.
  \item \textbf{Conditional harm:} “Your arm is broken; you cannot assist in combat until healed.”
  \item \textbf{Moral dilemmas:} The follower is offered a bribe to betray you; they resist but are shaken (Exposure +2).
  \item \textbf{Leave clues:} A missing follower leaves a trail the party can follow.
\end{itemize}

\section{Narrative Seeds}
\index{adventure seeds}\index{hooks!assets}\index{hooks!followers}
\paragraph{The Broken Ledger.}
Your \emph{Banking Charter} flags a forged letter-of-mark. Do you protect the account or the client's life?

\paragraph{The Bell at Dusk.}
A Mistlands \emph{Safehouse} stops answering the three-tone law. The bell rope's been cut from the outside.

\paragraph{The White Flag.}
Your Cap 3 mercenary refuses to take a fortified bridge without pay in hand. He's right—someone's playing both sides.

\paragraph{The Informant’s Debt.}
Your street informant calls in a favor from years ago. They need you to intimidate a landlord—or someone will die.

\section*{Quick Reference}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Assets \& Followers Quick Reference,fonttitle=\bfseries]
\textbf{Assets}
\begin{itemize}
  \item Cost/Time: Minor 4 XP/1 day; Standard 8 XP/1 week; Major 12 XP/1 month.
  \item Use: 1 free off-screen effect/session; 1 Boon to activate in-scene.
  \item Condition: Maintained \(\rightarrow\) Neglected (–1d) \(\rightarrow\) Compromised (repair in downtime).
  \item Portfolio: Maintain up to \textbf{Spirit} Assets without penalty.
\end{itemize}

\textbf{Followers}
\begin{itemize}
  \item Cost: Cap\(^2\) XP; Limit: \textbf{Presence} + 1 without strain.
  \item Assist: +\(\min(\)Cap, Skill\()\) dice, \textit{max +3d total from all sources}.
  \item Independent Action (party 1/scene): Scout \& Signal; Distract \& Draw; Fetch \& Carry; Cover \& Protect (cost: +Exposure or Harm 1).
  \item Tracks: Exposure/Harm 0–4; states mirror Assets.
  \item Loyalty: Wary (costs +1 XP upkeep), Steady, Devoted (once/arc downgrade complication).
\end{itemize}
\end{tcolorbox}

\section*{Closing Image}
\index{flourishes}
\textit{A shuttered lamp, a sealed ledger, a scarred gauntlet resting on a map. You don't move pieces—you invest in people and places. When you step into the street, the city already leans your way.}