% !TEX root = ../fates_edge_players_guide.tex

\chapter{Deck Generators}
\label{ch:deck-generators}

\begin{multicols}{2}

\index{generators!deck}\index{cards|see{generators, deck}}\index{procedures!random}\index{clocks}\index{Complication Points (CP)}\index{Boons}
In \textit{Fate’s Edge}, a well-cut deck can do the work of a crowded prep document. \textbf{Deck Generators} turn ordinary playing cards into places, people, pressures, and payoffs—then bind them to the core \emph{Clock}, \emph{CP}, and \emph{Boon} economy so your table’s surprises land with teeth.

\paragraph{Vignette: The Cut at Dusk.}
On the ferry stairs at Belworth, the GM fans four cards. A spade and a heart tap the boards: \emph{flooded causeway}, \emph{dike-guild reeve}. A club follows—\emph{audit at dawn}. Last, a diamond—\emph{priority ferry token}. The clock stands at eight, the wind at your backs, the line of eel-carts impatient. \emph{Play.}

\section{Core Structure}
\index{generators!structure}\index{suits}
Each suit maps to a narrative axis:

\begin{description}
  \item[{\large ♠} \textbf{Spades}]\index{Spades@\textbf{Spades}} Places: terrain, sites, hidden features, \emph{position} levers.
  \item[{\large ♡} \textbf{Hearts}]\index{Hearts@\textbf{Hearts}} People \& Factions: NPCs, circles, rivals, \emph{motives}.
  \item[{\large ♣} \textbf{Clubs}]\index{Clubs@\textbf{Clubs}} Complications \& Threats: hazards, pressure, \emph{setbacks}.
  \item[{\large ♢} \textbf{Diamonds}]\index{Diamonds@\textbf{Diamonds}} Rewards \& Leverage: treasure, writs, favors, \emph{advantages}.
\end{description}

\subsection*{Rank \(\rightarrow\) Clock}
\index{clocks!rank mapping}
The \emph{highest} rank among the drawn seed sets the primary Clock:
\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Rank} & \textbf{Clock Size} \\
\midrule
2–5 & 4 segments (\emph{Minor}) \\
6–10 & 6 segments (\emph{Standard}) \\
J,Q,K & 8 segments (\emph{Major}) \\
Ace & 10 segments (\emph{Pivotal}) \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Color Inflection}
\index{color (cards)}
\begin{itemize}
  \item \textbf{Black (♠♣)} skew toward travel hazards, tangible threats, fatigue.
  \item \textbf{Red (♡♢)} skew toward social pressure, reputation, moral leverage.
\end{itemize}

\section{Draw Procedures}
\index{procedures!draws}

\subsection*{Quick Hook (2 cards)}
Draw \(\mathbf{1\times♠}\) + \(\mathbf{1\times♡}\).\; The Spade sets the where; the Heart sets the who. Highest rank sets the main Clock. \emph{Optional:} If both are the same color, GM gains \textbf{+1 CP} for that opening beat. \index{CP economy}

\subsection*{Full Seed (4 cards)}
Draw until you have one of each suit: \(\spadesuit,\heartsuit,\clubsuit,\diamondsuit\).\; Map to \emph{Place, Actor, Complication, Leverage} in that order. Highest rank sets the mission Clock. Stash any extra Diamonds to foreshadow or \emph{cash} later for a scene-start \textbf{Boon}. \index{Boons}

\subsection*{Act Builder (3 cards/act)}
For each act/session, draw three cards: \emph{setting}, \emph{actor}, \emph{complication}. Keep any Diamonds aside as act payoffs. When the act Clock completes, flip the banked Diamond and pay it out in-fiction.

\section{Combo Rules}
\index{combos (cards)}\index{momentum}
If multiple drawn cards form poker-style patterns:
\begin{description}
  \item[\textbf{Pair} (same rank)] Recurring motif returns with a twist; reduce DV of \emph{follow-up} checks by 1 within that motif.
  \item[\textbf{Run} (3+ sequential ranks)] Momentum: \emph{reduce} the main Clock by 1 segment immediately.
  \item[\textbf{Flush} (3+ same suit)] Strong theme: the GM biases new clocks toward that axis (e.g., more place-driven obstacles).
  \item[\textbf{Face + Ace}] Reveal a hidden patron/power behind the drawn element; add a \emph{Shadow Clock} at 4 segments.
  \item[\textbf{All one color}] GM gains \textbf{+1 CP} at scene start.
\end{description}

\subsection*{Optional Jokers}
\index{Jokers}
Include 1–2 \emph{Jokers}. When drawn:
\begin{itemize}
  \item \textbf{Wild Place/Person:} Player names either the Spade or Heart entry this time.
  \item \textbf{Cut to Cost:} Everyone may take 1 \emph{Fatigue} to reshuffle and redraw the last card; if refused, GM gains +1 CP.
\end{itemize}

\section{Table Use: From Card to Fiction}
\index{GM tools}\index{framing}
\begin{enumerate}
  \item \textbf{Seed the Scene.} Read suits aloud; ask one question per card to the table (\emph{“What rumor makes this reeve uneasy?”}).
  \item \textbf{Set the Clock.} Size from rank; place it where players can see it.
  \item \textbf{Lean into Color.} Black = roads bite; Red = rooms breathe.
  \item \textbf{Spend the CP.} Use Clubs and black-weighted draws to justify early CP pressure.
  \item \textbf{Pay the Diamond.} Convert Diamonds into concrete leverage: writs, salvage tags, audience slots, sanctuary hours.
\end{enumerate}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Player Options with Cards,fonttitle=\bfseries]
\index{players!agency}\index{Boons}\index{Devil's Bargain}
\begin{itemize}
  \item \textbf{Boon for Bias:} Spend 1 \emph{Boon} before a pull to bias the next card toward a suit (draw two, keep one).
  \item \textbf{Cut for Consequence:} Take a GM-offered \emph{Bargain} to swap one revealed card with the top of the deck; the GM immediately marks +1 CP.
  \item \textbf{Keep the Diamond:} A player who \emph{earned} the Diamond describes how it appears on-screen when cashed.
\end{itemize}
\end{tcolorbox}

\section{Regional Deck Generators}
\index{world!regions}\index{generators!regional}
Each region carries its own texture. Use the same suit mapping; swap in entries below for local color. (Ranks 2–A correspond top-to-bottom.)

\subsection*{Viterra — \textit{The Last Kingdom}}
\index{Viterra}
\textbf{Spades — Places}
\begin{itemize}
  \item 2. Fen causeway stile with a toll-rod and a patient line of eel carts.
  \item 3. Hedgerow muster-green in the Dales; bows strung under apple trees.
  \item 4. Beacon hill above the Highlands; watch-fire grate and slate steps.
  \item 5. Belworth ferry-stairs with wet ledgers and a nervous horn.
  \item 6. Old iron-bloom quarry turned drill yard; hammer echoes carry.
  \item 7. Parish-stone maze where three maps disagree by a field.
  \item 8. Fairport tideworks at the river mouth; Dolmis swell under the planks.
  \item 9. Valora law quarter—archives, oath-rooms, Hall of Dawning nearby.
  \item 10. Tarlington counting fields by the mustering ground; quiet efficiency hums.
  \item J. River dike crown: ring of turf and timber; brotherhood bells for flood watch.
  \item Q. The Queen's Progress encampment—canvas palisade, fresh standards, full schedule.
  \item K. Hall of Dawning tilt-yard at first light; Dawn-knights run clean drills.
  \item A. Queen's Highway milestone above the Dolmis road; customs writ posted.
\end{itemize}

\textbf{Hearts — People and Factions}
\begin{itemize}
  \item 2. Fen reeve with a tally-rod; speaks for the dike guilds.
  \item 3. River-carter syndic who moves grain faster than rumor.
  \item 4. Parish surveyor with three maps and one opinion.
  \item 5. Quartermaster of the Dawn (logistics first, lances second).
  \item 6. Dales levy serjeant—longbow calm, cider breath.
  \item 7. Two-altars cleric-pair (Light circuit-preacher vs. Everflame canon lawyer).
  \item 8. Fairport shipwright with Dolmis cousins and a quiet skiff.
  \item 9. Fenwood comptroller who can conjure wagons with a signature.
  \item 10. Queen's Justiciar—law on the road, polite as a gallows.
  \item J. Border routier-captain who reads ledgers as well as ambushes.
  \item Q. The newly crowned queen—patient sums, sharp promises.
  \item K. Crown in Council (Fenwood, duchy envoys, guild voices).
  \item A. Tarling-blood rumor—a lost sigil surfaces; old loyalties twitch.
\end{itemize}

\textbf{Clubs — Complications \& Threats}
\begin{itemize}
  \item 2. Dike breach in black rain; carts bog and tempers sink.
  \item 3. Feast-day clash: Light vs. Everflame processions collide over tithes.
  \item 4. “Quiet tolls” sprout on the Highway; escorts sniff a trap.
  \item 5. Counting-house audit freezes your cargo until dawn.
  \item 6. Border-lace snarl: overlapping titles spark arrests mid-parish.
  \item 7. Isle refusal: Theona’s moot withholds levy; quay rumors harden.
  \item 8. Delta spat: Fairport vs. Marcott customs—barges stack three deep.
  \item 9. Routier arrears: free-company flips unless someone pays.
  \item 10. Salt pinch—import prices spike; bakers barricade.
  \item J. Dawn recall: your escort pulled to a flood-girded parish.
  \item Q. Aberielist intrigue: a royalist ring stirs against the crown.
  \item K. Levy call-up: dalesmen muster; wagons conscripted “for the realm”.
  \item A. Dolmis gale-train: bora winds slam the coast; schedules drown.
\end{itemize}

\textbf{Diamonds — Rewards \& Leverage}
\begin{itemize}
  \item 2. Ferry priority at a named Belworth crossing (once).
  \item 3. Dike-work allotment—brotherhood labor on your timetable.
  \item 4. Market-day license in Valora’s square.
  \item 5. Dawn escort letter (four lances at first light).
  \item 6. River-carter line—guaranteed haul on the grain artery.
  \item 7. Parish-map correction—move a border a hedgerow over.
  \item 8. Fairport customs seal for Dolmis-bound cargo.
  \item 9. County Thing ruling in your favor; local teeth, real bite.
  \item 10. Salt allotment from a guarded depot (winter only).
  \item J. Wardship of a minor fen-keep; men-at-arms “for now”.
  \item Q. Private audience with the queen; one secret exits as policy.
  \item K. Fenwood ducal warrant to enforce Highway customs.
  \item A. Coronation writ—temporary amnesty and tax remission for allies.
\end{itemize}

\subsection*{Acasia — \textit{Broken Marches}}
\index{Acasia}
\textbf{Spades — Places}
\begin{itemize}
  \item 2. Ruined signal tower over a contested valley.
  \item 3. Abandoned mine shaft with fresh chalk marks.
  \item 4. River ford where three clan ranges meet.
  \item 5. Burned steading reclaimed by bramble.
  \item 6. Standing stones that disagree with the new maps.
  \item 7. Pass-fort whose stones remember older oaths.
  \item 8. Hidden glen for clandestine parley.
  \item 9. Chieftain’s trophy hall; laughter with a knife in it.
  \item 10. Oath-fall where blood and water mix.
  \item J. Border watchtower with a loyal horn.
  \item Q. Hill market where iron buys silence.
  \item K. Strategic bridge that courts three banners.
  \item A. Lost ledger of lineages and feuds.
\end{itemize}

\textbf{Hearts — People and Factions}
\begin{itemize}
  \item 2. Young heir trying not to look afraid.
  \item 3. Veteran smuggler with clean boots.
  \item 4. Glen-seer reading crows and cloud breaks.
  \item 5. Consortium factor with clenched smiles.
  \item 6. Border captain with a pocket of broken seals.
  \item 7. Exiled noble rehearsing apologies.
  \item 8. Clan-mother who can count to war.
  \item 9. Company commander who keeps a fair ledger.
  \item 10. Reform preacher with a road-stained book.
  \item J. Master hunter tracking debts as well as deer.
  \item Q. Envoy who never says \emph{no}, only \emph{later}.
  \item K. Rebel banner-lord stitching counties together.
  \item A. Old spirit that remembers a wider forest.
\end{itemize}

\subsection*{Ecktoria — \textit{Marble \& Fire}}
\index{Ecktoria}
\textbf{Spades — Places}
\begin{itemize}
  \item 2. Public baths where deals rinse clean.
  \item 3. Temple precinct of quarreling choirs.
  \item 4. Magistrate’s court, docket humming.
  \item 5. Academy stacks with a sealed staircase.
  \item 6. Merchants’ floor where prices learn to dance.
  \item 7. Garrison parade, armor bright as law.
  \item 8. Cistern halls where whispers carry.
  \item 9. Garden villa with polite shadows.
  \item 10. Arena staging tomorrow’s heroics.
  \item J. Society chamber with masked rules.
  \item Q. Imperial archives—ink that bites back.
  \item K. Oracle’s cave, answers priced in shame.
  \item A. Senate floor under thunderous hands.
\end{itemize}

\subsection*{Ubral — \textit{Stone Between Spears}}
\index{Ubral}
\textbf{Spades — Places}
\begin{itemize}
  \item 2. Mountain shrine where breath shows vows.
  \item 3. Toll pass with counting scars on the gate.
  \item 4. Under-river where lanterns learn to float.
  \item 5. Nomad way-ring with smoke like script.
  \item 6. Fortified monastery deaf to bribes.
  \item 7. Crystal cave that hums to the wrong seasons.
  \item 8. Battlefield with orderly ghosts.
  \item 9. Caravan making the long turn home.
  \item 10. Oracle peak where the path tests knees.
  \item J. Hidden valley with inconvenient riches.
  \item Q. Distant outpost enforcing nearer laws.
  \item K. Tribal moot-stone with old grooves.
  \item A. Portal menhir that answers to weather.
\end{itemize}

\subsection*{Mistlands — \textit{Where Bells Speak}}
\index{Mistlands}
\textbf{Spades — Places}
\begin{itemize}
  \item 2. Causeway bell-frame missing its third tone.
  \item 3. Peat-cutters’ path that moves by night.
  \item 4. Fog orchard; fruit sweet, steps treacherous.
  \item 5. Fen barrow with polite bones.
  \item 6. Reed maze that echoes your name.
  \item 7. Ferry knoll with a rope cut from the outside.
  \item 8. Lantern-field where lights argue directions.
  \item 9. Sluice-gate that chooses favorites.
  \item 10. Bell-foundry that remembers voices.
  \item J. Embanked hall on stilts; law hangs from rafters.
  \item Q. Marsh-witch’s walk lined in driftwood charms.
  \item K. Tithe-island where debts get heavier.
  \item A. Greywater mirror: see what is owed.
\end{itemize}

\subsection*{Sun Coast — \textit{Ports and Pledges}}
\index{Sun Coast}
\textbf{Spades — Places}
\begin{itemize}
  \item 2. Vineyard switchback with lazy guards.
  \item 3. Drydock cradle creaking like old oaths.
  \item 4. Salt pan glinting with secrets.
  \item 5. Harbor chain that rises at dusk.
  \item 6. Customs quay with three stamps, one truth.
  \item 7. Pearl market where whispers are currency.
  \item 8. Lighthouse whose shadow points wrong.
  \item 9. Sea gate that eats small boats kindly.
  \item 10. Citrus court under painted awnings.
  \item J. Private quay flown under two flags.
  \item Q. Guild loggia tiled with confessions.
  \item K. Admiral’s balcony over polite threats.
  \item A. Stormline where ships earn names.
\end{itemize}

\section{Specialized Generators}
\index{generators!urban}\index{generators!wilderness}\index{generators!mystical}

\subsection*{Urban Encounter Generator}
\begin{description}
  \item[♠ Districts] markets, temples, slums, docks, courts, baths.
  \item[♡ City Folk] guards, brokers, criminals, priests, nobles.
  \item[♣ Dangers] corruption, gangs, fires, collapses, curfews.
  \item[♢ Opportunities] contacts, writs, maps, rumors, favors.
\end{description}
\emph{Tip:} On a \textbf{Flush} of Hearts, add a \emph{Reputation} clock; on \textbf{Flush} of Clubs, GM +1 CP.

\subsection*{Wilderness Journey Generator}
\begin{description}
  \item[♠ Terrain] rivers, ridges, thickets, ruins, passes.
  \item[♡ Encounters] rangers, nomads, hermits, spirits, caravans.
  \item[♣ Hazards] weather, predators, scarcity, getting lost.
  \item[♢ Discoveries] sanctuaries, shortcuts, forage, sign.
\end{description}
\emph{Tip:} On a \textbf{Run}, reduce the \emph{Travel Clock} by 1.

\subsection*{Mystical Adventure Generator}
\begin{description}
  \item[♠ Sites] leys, altars, libraries, portal stones, bell-precincts.
  \item[♡ Personae] mages, oathscribes, cultists, patrons, witnesses.
  \item[♣ Threats] backlash, taboos, rival rites, unquiet relics.
  \item[♢ Rewards] formulae, sigils, favors, sanctities, insights.
\end{description}
\emph{Tip:} \textbf{Face + Ace} creates a \emph{Patron Attention} clock at 4. \index{Patrons}

\section{Advanced Techniques}
\index{advanced!deck play}

\subsection*{Multi-Deck Integration}
Combine regional + specialized decks:
\begin{itemize}
  \item City-magic: Ecktoria (places) + Mystical (actors/complications).
  \item Border trek: Acasia (actors) + Wilderness (places/hazards).
  \item Sea intrigue: Sun Coast (places/rewards) + Urban (actors/complications).
\end{itemize}

\subsection*{Progressive Drawing}
\index{progress clocks!cards}
Start with a 2-card \emph{Hook}. Each time the main Clock crosses a quarter, draw one more card to introduce a new wrinkle. On completion, flip any banked Diamonds as epilogues.

\subsection*{Player-Involved Drawing}
\index{table procedures}\index{collaboration}
Let the player most invested in the scene draw. Offer a \emph{Boon-for-bias} before the draw or a \emph{Cut-for-consequence} after.

\section{Custom Deck Creation}
\index{homebrew!generators}
To homebrew a deck:
\begin{enumerate}
  \item Define theme/region focus and a one-line vibe.
  \item Brainstorm 13 entries per suit; keep three “quiet” ranks (2–4), five “working” (5–9), three “loud” (J–K), one “crown” (Ace).
  \item Seed mechanical hooks (Position shifts, DV nudges, CP hooks) into Clubs and Diamonds.
  \item Playtest: draw 10 random hands; adjust entries that repeat tone too often.
\end{enumerate}

\section*{Quick Reference}
\begin{tcolorbox}[colback=green!5!white,colframe=green!60!black,title=Deck Generators at a Glance,fonttitle=\bfseries]
\begin{itemize}
  \item \textbf{2-card Hook:} \(\spadesuit\) place + \(\heartsuit\) actor; highest rank = Clock.
  \item \textbf{4-card Seed:} add \(\clubsuit\) complication + \(\diamondsuit\) leverage.
  \item \textbf{Combos:} Pair = motif; Run = -1 segment; Flush = theme bias; Face+Ace = hidden power; one-color = GM +1 CP.
  \item \textbf{Boon-for-bias:} draw two, keep one suit.
  \item \textbf{Bank Diamonds:} foreshadow now, pay later.
\end{itemize}
\end{tcolorbox}

\section*{Closing Image}
\index{flourishes}
\textit{Cards slap the table; the ferry horn answers from the fog. The Clock ticks, the crowd leans, and somewhere a Diamond waits in a ledger with your name spelled right this time. Cut the deck. See who bleeds first.}

\end{multicols}
