\chapter{Enhanced Player Play}
\label{ch:enhanced-play}

% !TEX root = srd_main.tex
% SRD Insert: Optional Player-Character Death Rules

\section{Enhanced Coordination}

In Fate's Edge, teamwork isn't just about helping—it's about creating decisive moments through coordinated action. These rules help you and your allies combine efforts for greater impact.

\subsection{Core Coordination Principles}

\begin{itemize}
    \item \textbf{Fiction First:} Describe how you help before rolling dice
    \item \textbf{Clear Stakes:} State what help changes and what risk the helper accepts
    \item \textbf{One Spotlight:} Resolve one acting character's roll; fold assistance into that action
    \item \textbf{Visible Costs:} Story Beats, Fatigue, and consequences are tracked openly
\end{itemize}

\subsection{Basic Assistance}

\begin{itemize}
    \item \textbf{Declare Help:} An ally states a concrete contribution (tools, opening, cover)
    \item \textbf{Benefit:} Acting character gains +1d (up to table's assist cap)
    \item \textbf{Limits:} One assistant per PC per exchange by default
    \item \textbf{Cost:} Assistant accepts any risk named by the GM (SB, Fatigue, collateral)
\end{itemize}

\subsection{Cooperative Talents}

Certain talents enhance group coordination:

\textbf{Inspire (3 XP)} Once per scene, spend 1 Boon to choose one:
\begin{itemize}
    \item Bonded Ally: That ally gains +1 Boon and +1d on their next roll
    \item Self: Gain +1d on your next roll this scene  
    \item Rally: Each other PC in Near gains +1d on their next roll
    \item Tactical Coordination: All allies currently acting gain +1 Position
\end{itemize}

\textbf{Spotter's Mark (3 XP)} Aim a target (1 action). Until end of scene, allies in Near may claim +1d or +1 Effect once vs. that target.

\textbf{Battle Cant (2 XP)} Once per scene, establish silent signals. On next coordinated action where at least two PCs act, those PCs gain +1 Position.

\textbf{Shield Wall (4 XP)} If you and at least one ally each wield a shield and are adjacent: as a Defend action, grant +1d Defend to all in the Wall and convert first incoming Harm to Fatigue.

\subsection{Stacking Limits}

To prevent overwhelming combinations, a PC may benefit from at most two cooperative effects on the same action. Choose which apply when multiple sources could stack.

\subsection{Timing and Sequencing}

\begin{itemize}
    \item \textbf{Declare Order:} Players state intent in any order; resolve acting roll first
    \item \textbf{Ready/Overwatch:} Hold an action with clear trigger; resolve before next exchange
    \item \textbf{Refresh Windows:} "Once/scene" effects reset at scene end
\end{itemize}

\subsection{Working With Followers}

Followers can assist for up to +3d (or +4d with Exceptional Coordination talent). They can't receive PC-only benefits unless a talent states otherwise.

\subsection{Magic Coordination}

\begin{itemize}
    \item \textbf{Ritual Support:} Allies can assist Weave or Cast phases if fiction allows
    \item \textbf{Rite Preparation:} Helpers can provide components or maintain focus
    \item \textbf{Imbuement Synergy:} Multiple characters can prepare for shared magical effort
\end{itemize}

\subsection{Best Practices for Players}

\begin{enumerate}
    \item \textbf{Be Specific:} "I'll flank left to give you advantage" is better than "I help"
    \item \textbf{Accept Risk:} Good assistance often means putting yourself in harm's way
    \item \textbf{Think Creatively:} Tools, positioning, and timing matter more than raw dice
    \item \textbf{Support Bonds:} Use coordination to strengthen character relationships
    \item \textbf{Keep Momentum:} Quick, clear assistance keeps scenes moving
\end{enumerate}

Remember: The best coordination enhances the story and makes everyone's moments shine brighter together.

\section{Optional Player-Character Death}\label{sec:pc-death-optional}
\index{Death}\index{Last Stand}\index{Death Clock}\index{Patron's Claim}\index{Inheritance}\index{Resurrection}

By default, Fate's Edge treats death as rare and dramatic. Use any of the following modules (singly or in combination) to tune how lethal your table feels. All options respect \textbf{Story Beats (SB)} and \textbf{Obligation} as core currencies.

\subsection*{Baseline (Default)}\label{subsec:death-baseline}
PCs do not die on ordinary failures. Instead, they suffer \textbf{Harm}, Conditions, lost opportunities, or narrative costs. Death only occurs when a rule below is in play or the table agrees a scene warrants it.

\subsection*{Option A — Severe Harm Death}\label{subsec:death-severe}
\textbf{Trigger.} When a PC would take a \textbf{third} instance of Severe Harm (or escalate past the top of your harm track), they instead face death.\par
\textbf{Stave It Off.} The player may avoid death by choosing one: (1) mark \textbf{2 SB} \emph{and} take a \textsc{Maimed} permanent Condition, (2) accept a \textbf{Patron's Claim} (see Option D), or (3) convert the blow into a \textbf{Last Stand} (Option C).

\subsection*{Option B — Death Clock}\label{subsec:death-clock}
\textbf{Trigger.} Catastrophic consequences (falls, crushes, mortal wounds) fill a named \textbf{Death Clock} (4 or 6 ticks).\par
\textbf{While Ticking.} Actions that stabilize reduce the clock; taking further punishment advances it.\par
\textbf{When Full.} The character dies unless one of the following occurs immediately: spend \textbf{2 SB} to hold at full (buy a single action), accept a \textbf{Patron's Claim}, or another PC succeeds at \emph{Pull From the Brink} (risky, effect = clock –2 on success).

\subsection*{Option C — Last Stand}\label{subsec:last-stand}
\textbf{Trigger.} On lethal harm or a full Death Clock, the player may declare a \textbf{Last Stand}.\par
\textbf{Effect.} For the remainder of the scene, the PC acts with \textbf{+1 effect} and ignores new Harm. Each action automatically creates \textbf{1 SB}. When the scene ends, the character \textbf{dies} unless a \textbf{miracle} is secured (Patron rite, relic, or equivalent).

\subsection*{Option D — Patron's Claim}\label{subsec:patron-claim}
\textbf{Trigger.} On death, the PC's Patron (or a circling power) intervenes.\par
\textbf{Bargain.} The GM offers 1–3 \non-negotiable terms (e.g., \emph{Obligation +2}, \emph{forfeit a Gift}, \emph{become a vessel for a season}). If accepted, the PC lives; mark the costs immediately and record the \textbf{Claim} as an ongoing front. Refusal means the death proceeds.\par
\textbf{Note.} Claims should change the campaign; use sparingly.

\subsection*{Option E — Dramatic Exit & Inheritance}\label{subsec:inheritance}
\textbf{Dramatic Exit.} The player may choose a meaningful death that resolves a question or saves others.\par
\textbf{Inheritance.} The next PC created by that player inherits one of: (a) a \textbf{Relationship} (bond, contact, or rival), (b) a \textbf{Tool} (asset degraded one step), or (c) a \textbf{Lesson} (start with +1 XP toward a Talent used in the exit). Record how the world remembers them.

\subsection*{Option F — Return From Beyond}\label{subsec:return}
\textbf{Trigger.} A body, a name, and a path (rite, gate, bargain).\par
\textbf{Cost.} Treat as a High Rite with Obligation appropriate to the transgression. Set DV using the Rites system. On success, the PC returns \textbf{changed}: apply a \textsc{Scar} (permanent Condition) and \textbf{1 SB} to the rescuer. On failure, choose: lose the body, or return with a Patron's Claim.
\begin{tcolorbox}[title={Quick Picks},colback=gray!5,colframe=black]
\small \textbf{Low lethality:} Baseline + Death Clock only.\quad \textbf{Heroic tragedy:} Severe Harm Death + Last Stand + Inheritance.\quad \textbf{Dark bargains:} Any combo with Patron’s Claim.\end{tcolorbox}

\index{play!enhanced}\index{player-facing tools}\index{table culture}
The world of \textbf{Fate's Edge} sings when players help steer the tune. This chapter gathers \emph{player-facing dials}—lightweight, opt-in mechanics that turn your table's choices into momentum in the fiction of Viterra, Acasia, Ecktoria, Ubral, and the long coasts besides. Treat these as a menu: use a few, or layer many as your group grows comfortable.

\section{Player Resources}
\index{resources!player}\index{boons}\index{SB@Story Beats}

Two shared languages of play appear throughout this chapter:
\begin{description}
  \item[\textbf{Boons}] Core system edge; spend to power talents or convert to XP (per core rules).
  \item[\textbf{SB}] The GM's Story Beats; several options here invite players to accept SB in exchange for narrative benefits.
\end{description}

\section{Engagement Rewards}
\index{engagement}\index{investment tracker}

\subsection{Session Investment Tracker}
\index{investment tracker}
At session close, each player privately rates their \emph{investment} (1–3). The tracker rewards steady participation without penalizing quiet nights.

\begin{description}
  \item[\textbf{1 — Low}] You followed others' lead. Gain \textbf{+1d} once next session on any \emph{support} action.
  \item[\textbf{2 — Medium}] You took initiative in some scenes. Gain \textbf{+1d} once next session on a \emph{relationship} roll and bank \textbf{1 Momentum} (see \S\ref{sec:momentum-banking}).%
  \index{relationships}
  \item[\textbf{3 — High}] You anchored or elevated multiple scenes. Gain \textbf{+1d} twice next session (different scenes) and bank \textbf{1 Momentum}.%
\end{description}

\paragraph{Table Boon.} If \emph{all} players report 2+ in a session, the GM may award the table \textbf{1 free Boon} to assign at the start of the next session.

\subsection{Cultural Immersion Bonus}
\index{culture!immersion}\index{roleplay rewards}

Reward lived-in play with light, predictable benefits.
\begin{itemize}
  \item \textbf{Earn 1 Culture Point} when you enrich a scene with apt language, rites, or customs (Sea-Patter hail; bell-speech courtesy; Aeler craft-honorifics). Max \textbf{3/session}.%
  \item \textbf{Spend 3 Culture} for \textbf{+1d} on a culturally keyed roll.%
  \item \textbf{Spend 5 Culture} (across sessions) to declare a \emph{trusted door}: one institution in that culture treats you as Friendly for the next approach.%
  \index{factions!attitudes}
\end{itemize}

\section{Collaborative Play}
\index{collaborative play}\index{negotiation}\index{tables!contracts}

\subsection{Information Trading}
\index{information economy}

When the table hunts answers, convert curiosity into structure.

\paragraph{Request \& Price.}
State the question and choose a price the table accepts:
\begin{description}
  \item[\textbf{Devil's Bargain}] Offer the GM a future complication to "pay" for deeper intel.%
  \index{devil's bargain}
  \item[\textbf{Accept SB}] The GM gains \textbf{+1 SB} now; you gain a strategic clue or reduced DV.%
  \item[\textbf{Spend Boon}] Spend \textbf{1 Boon} to lower DV by 1 on the research/social approach.%
\end{description}

\paragraph{Creative Methods.}
Swap coin for color: poetry duels, shrine petitions, map-reading at a parish stone. If your method sings with setting, take \textbf{+1 effect} on the check.%
\index{effect}

\subsection{Clock Manipulation}
\index{clocks}\index{pacing}\label{sec:clock-manip}

Players can nudge tension without rewriting stakes.
\begin{itemize}
  \item \textbf{Slow} a visible Clock by \textbf{1 segment} by either \textbf{spending 1 Boon} \emph{or} \textbf{accepting +1 SB} into the scene (fiction must justify delay).
  \item \textbf{Hasten} a visible Clock by \textbf{1 segment} by \textbf{spending 1 Momentum} (\S\ref{sec:momentum-banking}) \emph{or} \textbf{taking a Devil's Bargain}. 
  \item \textbf{Limit:} Once per player per scene. If three or more players affect the same Clock in a scene, the GM gains \textbf{+1 SB}.
\end{itemize}

\subsection{Complication Bargaining}
\index{complications!player-asked}

Invite the kind of trouble you want to play.
\begin{itemize}
  \item Name a \emph{type}: social, physical, mystery, or moral (e.g., Valora court etiquette; Ubral scree; Isoka whispers; oath vs.\ mercy).
  \item The GM frames the complication accordingly and grants you \textbf{+1d} on your next roll \emph{within} that trouble, or reduces DV by 1 if your approach leans into the specified texture.
\end{itemize}

\section{Faction Awareness}
\index{factions}\index{attitudes}

\subsection{Loyalty Recognition}
\index{loyalty scale}
Track a simple ladder \((-3\ldots+3)\): \emph{Enemy, Hostile, Unfriendly, Neutral, Friendly, Supportive, Ally}. When you act \emph{with} an institution's aims, mark a \textbf{tick} toward the next step (GM pacing). When you betray a stated value, drop one step immediately and the GM banks \textbf{+1 SB} for future headaches.

\subsection{Cross-Cultural Synergies}
\index{synergy!cross-cultural}
Spotting a neat cultural combo (Zakov pilots + Kahfagia signals; Aeler engineering + Ecktoria charters) grants \textbf{+1d} once per scene the synergy is actively used.

\section{Advanced Techniques}
\index{advanced techniques}

\subsection{Momentum Banking}\label{sec:momentum-banking}
\index{momentum}
When your team resolves a Clock \emph{early} (segments unspent), bank \textbf{1 Momentum} per unused segment (max \textbf{2/session}). Spend 1 Momentum to:
\begin{itemize}
  \item Gain \textbf{+1d} on a future approach tied to that victory (\emph{lessons learned}); or
  \item \emph{Telescope} a travel beat (skip a routine obstacle the same route would present); or
  \item Trigger a \emph{Prepared Move}: declare a sensible minor setup you plausibly arranged off-screen.
\end{itemize}

\subsection{Escalation Management}
\index{escalation}
Trade heat for shape.
\begin{itemize}
  \item \textbf{De-escalate (spend 1 Boon):} Downgrade a Major consequence to Minor with a plausible concession (quiet tolls paid, harsh words eaten).
  \item \textbf{Redirect (accept +1 SB):} Shift pressure to a new venue or actor you name; the GM places that SB there as attention.
  \item \textbf{Truce (spend 1 Momentum):} Freeze a faction's hostility for one scene if you can cite a shared value (oath, rite, charter clause).%
\end{itemize}

\section{Character Creation Enhancements}
\index{character creation!enhancements}\index{bonds}\index{complications}

Players may take up to \textbf{2 Bonds} (+2 XP total) and up to \textbf{2 Starting Complications} (+4 XP total) for a cap of \textbf{34 XP}. Each unresolved starting Complication adds \textbf{+1 banked SB} to early scenes until cleared. Favor \emph{storyful} picks (clan honor, guild debt, patron notice) over pure math.

\section{Downtime}
\index{downtime}\index{bookkeeping}\index{prep}

Between sessions, tend your garden—quietly moving the world.

\subsection{Bookkeeping}
\index{downtime!bookkeeping}
\begin{itemize}
  \item Allocate XP (respecting days required).%
  \item Update Assets/Followers (status: Maintained, Neglected, Compromised).%
  \item Track Boons (conversion cap remains 2 XP/session).%
  \item Note SB Debt from unresolved Complications.%
\end{itemize}

\subsection{Activities}
\index{downtime!activities}
\begin{itemize}
  \item \textbf{Recovery:} Clear Harm/Exposure with scenes that show the work.%
  \item \textbf{Training:} Buy advances; narrate mentors, gyms, scriptoria.%
  \item \textbf{Research:} Lower DVs with good sources; log new leads.%
  \item \textbf{Social:} Strengthen Bonds; tune faction attitudes.%
  \item \textbf{Preparation:} Cache gear, sow rumors, line up ferries.%
\end{itemize}

\subsection{Strategic Considerations}
\index{downtime!strategy}
Clear SB-debt complications first; synchronize travel plans; decide which Asset gets love this interval, and which follower needs face-time to avoid Neglect.

\section{Implementation Timeline}
\index{onboarding}\index{session flow}

\subsection*{Sessions 1–3}
Investment Tracker, Information Trading (via Boons/Bargains), light Complication Bargaining.

\subsection*{Sessions 4–6}
Clock Manipulation (Boon/Momentum/SB), Cultural Immersion, Faction Awareness, Momentum Banking.

\subsection*{Sessions 7+}
Cross-Cultural Synergies, Escalation Management, advanced pacing choreography.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=At-Table Prompts,fonttitle=\bfseries]
\textbf{Seed the Scene.} "Ask for the kind of complication you want."\\
\textbf{Name the Stakes.} "Is this clock worth slowing? Who pays—Boon, SB, or Momentum?"\\
\textbf{Pay with Color.} "What custom or rite do you invoke to make this work?"\\
\textbf{Close the Loop.} "Mark your Investment; one sentence of what you learned."
\end{tcolorbox}

\subsection{War Mount (Asset)}
\label{subsec:war-mount}
\index{Assets!War Mount}
\index{Mounts!War Mount}

\textbf{Type:} Standard Asset \\
\textbf{XP Cost:} 8 XP \\
\textbf{Scale:} Personal \\
\textbf{Maintenance:} Regular care, feed, rest

A trained combat mount—warhorse, dune-strider, battle elk, ash lizard, or equivalent—conditioned for violence, chaos, and discipline.

\paragraph{Passive Benefits}
While mounted:
\begin{itemize}
  \item Increase overland travel speed and endurance.
  \item Ignore the first movement-based complication per session (mud, crowd, loose terrain).
  \item Mounted positioning may override terrain restrictions at GM discretion.
\end{itemize}

\paragraph{Mounted Combat}
If the rider has the \textbf{Cavalier} talent:
\begin{itemize}
  \item Melee attacks made while moving \textbf{Far → Near}: \textbf{+2d}
  \item Ranged attacks made while moving \textbf{Near → Far}: \textbf{+2d}
\end{itemize}

\paragraph{Boon Activation}
Spend \textbf{1 Boon} to invoke the mount dramatically:
\begin{itemize}
  \item Break through a hostile line or crowd
  \item Escape pursuit across hostile terrain
  \item Gain \textbf{Controlled Position} for a mounted action
  \item Advance or suppress a movement-related Clock
\end{itemize}

\paragraph{Risk and Wear}
On a \emph{Partial} or \emph{Miss} involving mounted actions, the GM may:
\begin{itemize}
  \item Mark \textbf{Mount Fatigue}
  \item Damage tack, barding, or terrain access
  \item Threaten or compromise the Asset
\end{itemize}

\paragraph{Mount Fatigue}
If Mount Fatigue reaches 3:
\begin{itemize}
  \item The mount cannot be used in combat until rested
  \item Further strain risks permanent injury or loss
\end{itemize}

\paragraph{Promoting a War Mount}
A War Mount may be promoted to a \textbf{Follower} if:
\begin{itemize}
  \item It is named and narratively significant
  \item The rider takes a dedicated Bond or Talent
\end{itemize}

As a Follower, the mount:
\begin{itemize}
  \item Can take actions independently (trample, defend, flee)
  \item Can suffer Harm
  \item May generate Boons through sacrifice or risk
\end{itemize}

\subsection*{Pike Discipline (Major Talent, 4 XP) \textnormal{[FORMATION] [CONTROL]}}
\index{Talents!Anti-Cavalry}

\textbf{Requirements:} Melee 2+, Discipline or Command 1+

You are trained to break mounted assaults through spacing, timing, and fearlessness.

\paragraph{Effect}
When you \textbf{Brace} against a mounted opponent:
\begin{itemize}
  \item Cancel all \textbf{Cavalier} movement bonuses
  \item Mounted attackers suffer \textbf{-1d} on the attack
\end{itemize}

\paragraph{On Success}
\begin{itemize}
  \item Reduce incoming Harm by 1
  \item You may force the rider to \textbf{Disengage} or risk mount injury
\end{itemize}

\paragraph{On Partial}
\begin{itemize}
  \item Reduce Harm by 1
  \item Mark 1 Fatigue or lose formation
\end{itemize}

\paragraph{Limits}
\begin{itemize}
  \item Requires braced footing or prepared terrain
  \item Ineffective if flanked, surprised, or outnumbered
\end{itemize}

\newpage
\section{Between Sessions Activities Log}
\index{logs!between sessions}\index{play aids}

Use or print the following trackers.

\subsection{Character Advancement}
\index{logs!advancement}
\subsubsection*{XP Allocation}
\begin{itemize}
  \item \textbf{Total XP Available}: \_\_\_\_\_
  \item \textbf{Attributes Spent}: \_\_\_\_\_ (=\_\_\_ days)
  \item \textbf{Skills Spent}: \_\_\_\_\_ (=\_\_\_ days)
  \item \textbf{Remaining XP}: \_\_\_\_\_
\end{itemize}

\subsubsection*{Attribute Improvements}
\begin{longtable}{|p{3cm}|c|c|c|c|}
\hline
\textbf{Attribute} & \textbf{Old} & \textbf{New} & \textbf{Cost} & \textbf{Days} \\
\hline
Body & & & & \\
Spirit & & & & \\
Presence & & & & \\
Wits & & & & \\
\hline
\end{longtable}

\subsubsection*{Skill Improvements}
\begin{longtable}{|p{4cm}|c|c|c|c|}
\hline
\textbf{Skill} & \textbf{Old} & \textbf{New} & \textbf{Cost} & \textbf{Days} \\
\hline
Arcana & & & & \\
Combat & & & & \\
Investigate & & & & \\
Lore & & & & \\
Move & & & & \\
Notice & & & & \\
Physique & & & & \\
Resolve & & & & \\
Stealth & & & & \\
Sway & & & & \\
Survival & & & & \\
\hline
\end{longtable}

\subsection{Asset \& Follower Management}
\index{logs!assets}\index{logs!followers}
\subsubsection*{Assets}
\begin{longtable}{|p{4cm}|p{2cm}|p{2cm}|p{2.2cm}|}
\hline
\textbf{Name} & \textbf{Tier} & \textbf{Status} & \textbf{Notes} \\
\hline
 & & & \\
 & & & \\
\hline
\end{longtable}

\subsubsection*{Followers}
\begin{longtable}{|p{3.2cm}|p{2cm}|p{1.4cm}|p{1.6cm}|p{2cm}|}
\hline
\textbf{Name} & \textbf{Role} & \textbf{Harm} & \textbf{Exposure} & \textbf{Status} \\
\hline
 & & & & \\
 & & & & \\
\hline
\end{longtable}

\subsection{Bonds \& Complications}
\index{logs!bonds}\index{logs!complications}
\subsubsection*{Bonds Updated}
\begin{itemize}
  \item \textbf{With:} \_\_\_\_\_ — \textbf{Change:} \_\_\_\_\_
\end{itemize}
\subsubsection*{Complications}
\begin{longtable}{|p{5cm}|p{5cm}|}
\hline
\textbf{Complication} & \textbf{Resolution or Status} \\
\hline
 &  \\
 &  \\
\hline
\end{longtable}

\subsection{Boons \& Momentum Summary}
\index{logs!boons}\index{logs!momentum}
\begin{itemize}
  \item \textbf{Boons Held:} \_\_/5 \quad \textbf{Converted to XP}: \_\_ (max 2 XP/session)
  \item \textbf{Momentum Banked:} \_\_ (max 2/session)
\end{itemize}

\subsection{Campaign Clocks}
\index{campaign clocks}
\begin{itemize}
  \item \textbf{Mandate:} \_\_ / 6 \qquad \textbf{Crisis:} \_\_ / 6
  \item \textbf{Notables:} \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\end{itemize}