% !TEX root = ../fates_edge_players_guide.tex

\chapter{Example Character Builds}
\label{ch:example-builds}

\begin{multicols}{2}

\index{builds|see{characters, example}}\index{pregens}\index{advancement!paths}
In \textit{Fate’s Edge}, a build isn’t just math—it’s a point of view. These ten ready-to-play characters marry crisp mechanics to living voices. Use them as pregens, templates, or lodestars for your own designs.

\section*{How to Read These Builds}
\index{Character builds!format}
Each entry includes \textbf{Concept}, a \textbf{Build Summary} (attributes, skills, talent), typical \textbf{Dice Pools}, a \textbf{Signature Move}, \textbf{Complication Hooks}, and a sketched \textbf{Advancement Path}.

\paragraph{About XP math.}\index{XP!starting} Unless noted, each build spends \textbf{29 XP} (with \textbf{1 XP unspent}) using a consistent starting template:
\begin{itemize}
  \item \textbf{Attributes:} one at 3 (9 XP), one at 2 (6 XP), two at 1 (0 XP) \(\Rightarrow 15\) XP
  \item \textbf{Skills:} one at 3 (6 XP), one at 2 (4 XP) \(\Rightarrow 10\) XP
  \item \textbf{Talent:} one starting talent (4 XP)
  \item \textbf{Total:} 29 XP (\( +1\) XP unspent). Add Bonds/Complications as your table permits.
\end{itemize}

\section{1. The Witch Hunter}
\index{Character builds!Witch Hunter}\index{Witch Hunter}
\textit{“I was meant for lecture halls. The halls spat me out. The roads taught me the rest.”}

\textbf{Concept:} A Thepyrgosi-trained washout who now hunts illicit rites, selling dangerous finds to the University and sleeping badly afterwards.

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Wits 3 (9), Body 2 (6), Spirit 1 (0), Presence 1 (0)
  \item \textbf{Skills:} Skulduggery 3 (6), Sway 2 (4)
  \item \textbf{Talent:} Witch-Finder’s Sigil (4) — you can \emph{test} for recent ritework; on a clean success, gain a Boon to the next related action.
\end{itemize}
\textbf{Dice Pools:}
\begin{itemize}
  \item Wits + Skulduggery (locks, traps, contraband): \textbf{6d}
  \item Presence + Sway (press a suspect): \textbf{3d} (4d with positioning/Boons)
\end{itemize}
\textbf{Signature Move:} \emph{Cold Reading} — Once per scene, ask the GM one yes/no about a suspect’s \emph{recent} ritual activity after a brief interaction.

\textbf{Complication Hooks:}
\begin{itemize}
  \item The University’s purse strings knot at awkward moments.
  \item Evidence chains fray; informants burn.
  \item You know the words to open doors you should leave shut.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item \textit{Early:} Lore 2 (4) or Investigate 2 (4) for paper trails.
  \item \textit{Mid:} Resolve 2 (4) and the \emph{Oath-Tongue} talent (6) to bind deals.
  \item \textit{Late:} Presence 2 (6) and \emph{Network Builder} (6) to run informants safely (safer).
\end{itemize}

\section{2. The Oath-Bound Blade}
\index{Character builds!Oath-Bound Blade}
\textit{“Name the charge. I will be the wall.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Body 3 (9), Spirit 2 (6), Wits 1, Presence 1
  \item \textbf{Skills:} Melee 3 (6), Resolve 2 (4)
  \item \textbf{Talent:} Vow-Keeper’s Word (4) — when acting in defense of a sworn charge, begin from \emph{Controlled} once per scene.
\end{itemize}
\textbf{Dice Pools:} Body+Melee \textbf{6d}; Spirit+Resolve \textbf{5d}.\\
\textbf{Signature Move:} \emph{By My Word}.\\
\textbf{Hooks:} Oaths tangle; mercy and mandate collide.\\
\textbf{Advancement:} Shield Mastery (6); Presence 2 (6) for command.

\section{3. The Spirit-Speaker}
\index{Character builds!Spirit-Speaker}
\textit{“Ask. Pay. Thank. Repeat.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Spirit 3 (9), Wits 2 (6), Body 1, Presence 1
  \item \textbf{Skills:} Arcana 3 (6), Sway 2 (4)
  \item \textbf{Talent:} Spirit Whisperer (4) — once per scene gain a short, truthful impression from a willing local spirit.
\end{itemize}
\textbf{Dice Pools:} Spirit+Arcana \textbf{6d}; Spirit+Sway \textbf{5d}.\\
\textbf{Signature Move:} \emph{Whisper to the Unseen}.\\
\textbf{Hooks:} Favors accrue; taboos tighten.\\
\textbf{Advancement:} Familiar (6); Arcana 4 (8).

\section{4. The Border-Warden}
\index{Character builds!Border-Warden}
\textit{“Lines on maps mean little. Lines in mud mean more.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Wits 3 (9), Body 2 (6), Spirit 1, Presence 1
  \item \textbf{Skills:} Survival 3 (6), Stealth 2 (4)
  \item \textbf{Talent:} Territory Sense (4) — in familiar ground, \emph{Notice/Survey} starts one step better in Position once per scene.
\end{itemize}
\textbf{Dice Pools:} Wits+Survival \textbf{6d}; Body+Stealth \textbf{5d}.\\
\textbf{Signature Move:} \emph{Warden’s Eye}.\\
\textbf{Hooks:} Patrol politics; shifting posts.\\
\textbf{Advancement:} Safehouse (4); Survival 4 (8).

\section{5. The Guild Agent}
\index{Character builds!Guild Agent}
\textit{“A stamped paper is a door. I bring keys.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Presence 3 (9), Wits 2 (6), Body 1, Spirit 1
  \item \textbf{Skills:} Subterfuge 3 (6), Investigate 2 (4)
  \item \textbf{Talent:} Guild Connections (4) — once/session invoke guild backing for \emph{Controlled} entry to a civic negotiation.
\end{itemize}
\textbf{Dice Pools:} Presence+Subterfuge \textbf{6d}; Wits+Investigate \textbf{5d}.\\
\textbf{Signature Move:} \emph{Guild Authority}.\\
\textbf{Hooks:} Rival charters; audit knives.\\
\textbf{Advancement:} Contact Network (4); Subterfuge 4 (8).

\section{6. The Spirit-Touched Warrior}
\index{Character builds!Spirit-Touched Warrior}
\textit{“An old hand steadies mine.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Body 3 (9), Spirit 2 (6), Wits 1, Presence 1
  \item \textbf{Skills:} Melee 3 (6), Survival 2 (4)
  \item \textbf{Talent:} Ancestral Echo (4) — once/combat, spend a Boon to add +2d to a strike if you name the forebear who guides it.
\end{itemize}
\textbf{Dice Pools:} Body+Melee \textbf{6d}; Spirit+Survival \textbf{5d}.\\
\textbf{Signature Move:} \emph{Echo Strike}.\\
\textbf{Hooks:} Family debts; omen dreams.\\
\textbf{Advancement:} Spirit-Weapon (6); Body 4 (12).

\section{7. The Lore-Keeper}
\index{Character builds!Lore-Keeper}
\textit{“Ink can be a weapon if you read the edge.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Wits 3 (9), Spirit 2 (6), Body 1, Presence 1
  \item \textbf{Skills:} Lore 3 (6), Arcana 2 (4)
  \item \textbf{Talent:} Scholarly Insight (4) — when you present sourced citations, lower DV by 1 for the relevant argument or analysis.
\end{itemize}
\textbf{Dice Pools:} Wits+Lore \textbf{6d}; Wits+Arcana \textbf{5d}.\\
\textbf{Signature Move:} \emph{Hidden Connection}.\\
\textbf{Hooks:} Gatekept archives; rival theses.\\
\textbf{Advancement:} Library (4); Lore 4 (8).

\section{8. The Cycle-Warden}
\index{Character builds!Cycle-Warden}
\textit{“All growth is theft from somewhere. Pay back.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Spirit 3 (9), Wits 2 (6), Body 1, Presence 1
  \item \textbf{Skills:} Survival 3 (6), Lore 2 (4)
  \item \textbf{Talent:} Balance Sense (4) — when you restore, name a cost; if paid in-scene, gain a Boon immediately.
\end{itemize}
\textbf{Dice Pools:} Spirit+Survival \textbf{6d}; Wits+Lore \textbf{5d}.\\
\textbf{Signature Move:} \emph{Cycle’s Price}.\\
\textbf{Hooks:} Industry vs. watershed.\\
\textbf{Advancement:} Grove Sanctuary (4); Survival 4 (8).

\section{9. The Story-Weaver}
\index{Character builds!Story-Weaver}
\textit{“Tell it right and even the room must listen.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Presence 3 (9), Wits 2 (6), Body 1, Spirit 1
  \item \textbf{Skills:} Performance 3 (6), Insight 2 (4)
  \item \textbf{Talent:} Narrative Influence (4) — once/session reframe a social scene; on a clean success, shift one attitude step.
\end{itemize}
\textbf{Dice Pools:} Presence+Performance \textbf{6d}; Wits+Insight \textbf{5d}.\\
\textbf{Signature Move:} \emph{Story’s Turn}.\\
\textbf{Hooks:} Counter-bards; public gaffes.\\
\textbf{Advancement:} Venue (4); Performance 4 (8).

\section{10. The Focused Adept}
\index{Character builds!Focused Adept}
\textit{“Breath is a blade you sharpen from the inside.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Spirit 3 (9), Body 2 (6), Wits 1, Presence 1
  \item \textbf{Skills:} Resolve 3 (6), Melee 2 (4)
  \item \textbf{Talent:} Mind–Body Unity (4) — spend a Boon to ignore the first point of Fatigue/Light Harm this scene.
\end{itemize}
\textbf{Dice Pools:} Spirit+Resolve \textbf{6d}; Body+Melee \textbf{5d}.\\
\textbf{Signature Move:} \emph{Focused Flow}.\\
\textbf{Hooks:} Austerity’s costs; brittle calm.\\
\textbf{Advancement:} Spirit 4 (12); Meditation Sanctuary (4).

\section{11. The Faithful Petitioner}
\index{Character builds!Faithful Petitioner}
\textit{“Mercy is a kind of architecture.”}

\textbf{Build Summary (29/30 XP):}
\begin{itemize}
  \item \textbf{Attributes:} Spirit 3 (9), Presence 2 (6), Body 1, Wits 1
  \item \textbf{Skills:} Faith 3 (6), Sway 2 (4)
  \item \textbf{Talent:} Divine Interpretation (4) — once/session, treat a fresh, plausible omen as a Boon for a guidance action.
\end{itemize}
\textbf{Dice Pools:} Spirit+Faith \textbf{6d}; Presence+Sway \textbf{5d}.\\
\textbf{Signature Move:} \emph{Divine Sign}.\\
\textbf{Hooks:} Schisms; public doubt.\\
\textbf{Advancement:} Shrine (4); Faith 4 (8).

\begin{tcolorbox}[colback=green!5!white,colframe=green!60!black,title=Quick Conversion, fonttitle=\bfseries]
\begin{itemize}
  \item Want tougher starts? Shift the unspent 1 XP into \emph{savings} and take a Bond for +2 XP to buy a third skill at 2.
  \item Want louder talents? Bank your early attribute raise and pick a 6 XP talent as your first advancement beat.
  \item Want gear leverage? Trade the next skill-up for a Minor Asset (4 XP) tied to your concept.
\end{itemize}
\end{tcolorbox}

\end{multicols}
