% !TEX root = ../fates_edge_players_guide.tex

\chapter{Enhanced Player Play}
\label{ch:enhanced-play}

\begin{multicols}{2}

\index{play!enhanced}\index{player-facing tools}\index{table culture}
The world of \textbf{Fate’s Edge} sings when players help steer the tune. This chapter gathers \emph{player-facing dials}—lightweight, opt-in mechanics that turn your table’s choices into momentum in the fiction of Viterra, Acasia, Ecktoria, Ubral, and the long coasts besides. Treat these as a menu: use a few, or layer many as your group grows comfortable.

\section{Player Resources}
\index{resources!player}\index{boons}\index{SB@Story Beats}

Two shared languages of play appear throughout this chapter:
\begin{description}
  \item[\textbf{Boons}] Core system edge; spend to power talents or convert to XP (per core rules).
  \item[\textbf{SB}] The GM’s Story Beats; several options here invite players to accept SB in exchange for narrative benefits.
\end{description}

\section{Engagement Rewards}
\index{engagement}\index{investment tracker}

\subsection{Session Investment Tracker}
\index{investment tracker}
At session close, each player privately rates their \emph{investment} (1–3). The tracker rewards steady participation without penalizing quiet nights.

\begin{description}
  \item[\textbf{1 — Low}] You followed others’ lead. Gain \textbf{+1d} once next session on any \emph{support} action.
  \item[\textbf{2 — Medium}] You took initiative in some scenes. Gain \textbf{+1d} once next session on a \emph{relationship} roll and bank \textbf{1 Momentum} (see \S\ref{sec:momentum-banking}).%
  \index{relationships}
  \item[\textbf{3 — High}] You anchored or elevated multiple scenes. Gain \textbf{+1d} twice next session (different scenes) and bank \textbf{1 Momentum}.%
\end{description}

\paragraph{Table Boon.} If \emph{all} players report 2+ in a session, the GM may award the table \textbf{1 free Boon} to assign at the start of the next session.

\subsection{Cultural Immersion Bonus}
\index{culture!immersion}\index{roleplay rewards}

Reward lived-in play with light, predictable benefits.
\begin{itemize}
  \item \textbf{Earn 1 Culture Point} when you enrich a scene with apt language, rites, or customs (Sea-Patter hail; bell-speech courtesy; Aeler craft-honorifics). Max \textbf{3/session}.%
  \item \textbf{Spend 3 Culture} for \textbf{+1d} on a culturally keyed roll.%
  \item \textbf{Spend 5 Culture} (across sessions) to declare a \emph{trusted door}: one institution in that culture treats you as Friendly for the next approach.%
  \index{factions!attitudes}
\end{itemize}

\section{Collaborative Play}
\index{collaborative play}\index{negotiation}\index{tables!contracts}

\subsection{Information Trading}
\index{information economy}

When the table hunts answers, convert curiosity into structure.

\paragraph{Request \& Price.}
State the question and choose a price the table accepts:
\begin{description}
  \item[\textbf{Devil’s Bargain}] Offer the GM a future complication to “pay” for deeper intel.%
  \index{devil's bargain}
  \item[\textbf{Accept SB}] The GM gains \textbf{+1 SB} now; you gain a strategic clue or reduced DV.%
  \item[\textbf{Spend Boon}] Spend \textbf{1 Boon} to lower DV by 1 on the research/social approach.%
\end{description}

\paragraph{Creative Methods.}
Swap coin for color: poetry duels, shrine petitions, map-reading at a parish stone. If your method sings with setting, take \textbf{+1 effect} on the check.%
\index{effect}

\subsection{Clock Manipulation}
\index{clocks}\index{pacing}\label{sec:clock-manip}

Players can nudge tension without rewriting stakes.
\begin{itemize}
  \item \textbf{Slow} a visible Clock by \textbf{1 segment} by either \textbf{spending 1 Boon} \emph{or} \textbf{accepting +1 SB} into the scene (fiction must justify delay).
  \item \textbf{Hasten} a visible Clock by \textbf{1 segment} by \textbf{spending 1 Momentum} (\S\ref{sec:momentum-banking}) \emph{or} \textbf{taking a Devil’s Bargain}. 
  \item \textbf{Limit:} Once per player per scene. If three or more players affect the same Clock in a scene, the GM gains \textbf{+1 SB}.
\end{itemize}

\subsection{Complication Bargaining}
\index{complications!player-asked}

Invite the kind of trouble you want to play.
\begin{itemize}
  \item Name a \emph{type}: social, physical, mystery, or moral (e.g., Valora court etiquette; Ubral scree; Isoka whispers; oath vs.\ mercy).
  \item The GM frames the complication accordingly and grants you \textbf{+1d} on your next roll \emph{within} that trouble, or reduces DV by 1 if your approach leans into the specified texture.
\end{itemize}

\section{Faction Awareness}
\index{factions}\index{attitudes}

\subsection{Loyalty Recognition}
\index{loyalty scale}
Track a simple ladder \((-3\ldots+3)\): \emph{Enemy, Hostile, Unfriendly, Neutral, Friendly, Supportive, Ally}. When you act \emph{with} an institution’s aims, mark a \textbf{tick} toward the next step (GM pacing). When you betray a stated value, drop one step immediately and the GM banks \textbf{+1 SB} for future headaches.

\subsection{Cross-Cultural Synergies}
\index{synergy!cross-cultural}
Spotting a neat cultural combo (Zakov pilots + Kahfagia signals; Aeler engineering + Ecktoria charters) grants \textbf{+1d} once per scene the synergy is actively used.

\section{Advanced Techniques}
\index{advanced techniques}

\subsection{Momentum Banking}\label{sec:momentum-banking}
\index{momentum}
When your team resolves a Clock \emph{early} (segments unspent), bank \textbf{1 Momentum} per unused segment (max \textbf{2/session}). Spend 1 Momentum to:
\begin{itemize}
  \item Gain \textbf{+1d} on a future approach tied to that victory (\emph{lessons learned}); or
  \item \emph{Telescope} a travel beat (skip a routine obstacle the same route would present); or
  \item Trigger a \emph{Prepared Move}: declare a sensible minor setup you plausibly arranged off-screen.
\end{itemize}

\subsection{Escalation Management}
\index{escalation}
Trade heat for shape.
\begin{itemize}
  \item \textbf{De-escalate (spend 1 Boon):} Downgrade a Major consequence to Minor with a plausible concession (quiet tolls paid, harsh words eaten).
  \item \textbf{Redirect (accept +1 SB):} Shift pressure to a new venue or actor you name; the GM places that SB there as attention.
  \item \textbf{Truce (spend 1 Momentum):} Freeze a faction’s hostility for one scene if you can cite a shared value (oath, rite, charter clause).%
\end{itemize}

\section{Character Creation Enhancements}
\index{character creation!enhancements}\index{bonds}\index{complications}

Players may take up to \textbf{2 Bonds} (+2 XP total) and up to \textbf{2 Starting Complications} (+4 XP total) for a cap of \textbf{34 XP}. Each unresolved starting Complication adds \textbf{+1 banked SB} to early scenes until cleared. Favor \emph{storyful} picks (clan honor, guild debt, patron notice) over pure math.

\section{Downtime}
\index{downtime}\index{bookkeeping}\index{prep}

Between sessions, tend your garden—quietly moving the world.

\subsection{Bookkeeping}
\index{downtime!bookkeeping}
\begin{itemize}
  \item Allocate XP (respecting days required).%
  \item Update Assets/Followers (status: Maintained, Neglected, Compromised).%
  \item Track Boons (conversion cap remains 2 XP/session).%
  \item Note SB Debt from unresolved Complications.%
\end{itemize}

\subsection{Activities}
\index{downtime!activities}
\begin{itemize}
  \item \textbf{Recovery:} Clear Harm/Exposure with scenes that show the work.%
  \item \textbf{Training:} Buy advances; narrate mentors, gyms, scriptoria.%
  \item \textbf{Research:} Lower DVs with good sources; log new leads.%
  \item \textbf{Social:} Strengthen Bonds; tune faction attitudes.%
  \item \textbf{Preparation:} Cache gear, sow rumors, line up ferries.%
\end{itemize}

\subsection{Strategic Considerations}
\index{downtime!strategy}
Clear SB-debt complications first; synchronize travel plans; decide which Asset gets love this interval, and which follower needs face-time to avoid Neglect.

\section{Implementation Timeline}
\index{onboarding}\index{session flow}

\subsection*{Sessions 1–3}
Investment Tracker, Information Trading (via Boons/Bargains), light Complication Bargaining.

\subsection*{Sessions 4–6}
Clock Manipulation (Boon/Momentum/SB), Cultural Immersion, Faction Awareness, Momentum Banking.

\subsection*{Sessions 7+}
Cross-Cultural Synergies, Escalation Management, advanced pacing choreography.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=At-Table Prompts,fonttitle=\bfseries]
\textbf{Seed the Scene.} “Ask for the kind of complication you want.”\\
\textbf{Name the Stakes.} “Is this clock worth slowing? Who pays—Boon, SB, or Momentum?”\\
\textbf{Pay with Color.} “What custom or rite do you invoke to make this work?”\\
\textbf{Close the Loop.} “Mark your Investment; one sentence of what you learned.”
\end{tcolorbox}

\section{Between Sessions Activities Log}
\index{logs!between sessions}\index{play aids}

Use or print the following trackers.

\subsection{Character Advancement}
\index{logs!advancement}
\subsubsection*{XP Allocation}
\begin{itemize}
  \item \textbf{Total XP Available}: \_\_\_\_\_
  \item \textbf{Attributes Spent}: \_\_\_\_\_ (=\_\_\_ days)
  \item \textbf{Skills Spent}: \_\_\_\_\_ (=\_\_\_ days)
  \item \textbf{Remaining XP}: \_\_\_\_\_
\end{itemize}

\subsubsection*{Attribute Improvements}
\begin{tabular}{|p{3cm}|c|c|c|c|}
\hline
\textbf{Attribute} & \textbf{Old} & \textbf{New} & \textbf{Cost} & \textbf{Days} \\
\hline
Body & & & & \\
Spirit & & & & \\
Presence & & & & \\
Wits & & & & \\
\hline
\end{tabular}

\subsubsection*{Skill Improvements}
\begin{tabular}{|p{4cm}|c|c|c|c|}
\hline
\textbf{Skill} & \textbf{Old} & \textbf{New} & \textbf{Cost} & \textbf{Days} \\
\hline
Arcana & & & & \\
Combat & & & & \\
Investigate & & & & \\
Lore & & & & \\
Move & & & & \\
Notice & & & & \\
Physique & & & & \\
Resolve & & & & \\
Stealth & & & & \\
Sway & & & & \\
Survival & & & & \\
\hline
\end{tabular}

\subsection{Asset \& Follower Management}
\index{logs!assets}\index{logs!followers}
\subsubsection*{Assets}
\begin{tabular}{|p{4cm}|p{2cm}|p{2cm}|p{2.2cm}|}
\hline
\textbf{Name} & \textbf{Tier} & \textbf{Status} & \textbf{Notes} \\
\hline
 & & & \\
 & & & \\
\hline
\end{tabular}

\subsubsection*{Followers}
\begin{tabular}{|p{3.2cm}|p{2cm}|p{1.4cm}|p{1.6cm}|p{2cm}|}
\hline
\textbf{Name} & \textbf{Role} & \textbf{Harm} & \textbf{Exposure} & \textbf{Status} \\
\hline
 & & & & \\
 & & & & \\
\hline
\end{tabular}

\subsection{Bonds \& Complications}
\index{logs!bonds}\index{logs!complications}
\subsubsection*{Bonds Updated}
\begin{itemize}
  \item \textbf{With:} \_\_\_\_\_ — \textbf{Change:} \_\_\_\_\_
\end{itemize}
\subsubsection*{Complications}
\begin{tabular}{|p{5cm}|p{5cm}|}
\hline
\textbf{Complication} & \textbf{Resolution or Status} \\
\hline
 &  \\
 &  \\
\hline
\end{tabular}

\subsection{Boons \& Momentum Summary}
\index{logs!boons}\index{logs!momentum}
\begin{itemize}
  \item \textbf{Boons Held:} \_\_/5 \quad \textbf{Converted to XP}: \_\_ (max 2 XP/session)
  \item \textbf{Momentum Banked:} \_\_ (max 2/session)
\end{itemize}

\subsection{Campaign Clocks}
\index{campaign clocks}
\begin{itemize}
  \item \textbf{Mandate:} \_\_ / 6 \qquad \textbf{Crisis:} \_\_ / 6
  \item \textbf{Notables:} \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\end{itemize}

\subsection{Notes \& Reflections}
\index{journals}
\vspace{2.5cm}

\end{multicols}
