\chapter{Character Advancement and Capabilities}
\label{ch:advancement}

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Philosophy of Growth, fonttitle=\bfseries]
\index{advancement!philosophy}
\textbf{Every XP spent is a choice about who your character becomes -- invest in yourself, your network, or your legacy.}

Your character is not just a collection of numbers; they are a living story. When you spend XP on an Attribute, you are saying ``I become stronger, smarter, more resilient.'' When you invest in a Follower or an Asset, you are saying ``I build relationships and influence that outlast me.'' When you buy a Talent, you are saying ``This is my signature, my unique edge.''

\medskip
\noindent There is no single ``correct'' way to grow. A lone duelist (see \ref{sec:player-archetypes}) might pour everything into personal skill; a merchant prince might build a trading empire. Both are valid. The question is: \textit{what kind of story do you want to tell?}
\end{tcolorbox}

\section*{Why This Chapter Matters}
\index{advancement}\index{Experience Points (XP)}

In this game, growth is not just about numbers --- it is about deciding who your
character becomes. Advancement through \textbf{Experience Points (XP)} lets you
shape your capabilities, influence, and legacy in the world. Every XP choice is
a statement about your character's priorities and the mark they leave behind.

Your capabilities are built on four core \textbf{Attributes} and a set of
specialized \textbf{Skills}. This chapter explains how they work together,
what advancement looks like in play, and offers player‑facing tips for making
satisfying choices.

% ============================================================
% CORE ATTRIBUTES
% ============================================================
\section{Core Attributes}
\label{sec:core-attributes}
\index{Attributes}\index{Body}\index{Wits}\index{Spirit}\index{Presence}

Attributes represent your character's fundamental capabilities. Each is rated
from 1 to 5; higher numbers mean greater potential in that area.

\subsection{Body}
\label{sec:attribute-body}
\index{Body}
Physical strength, endurance, coordination, and health.
\begin{itemize}
  \item \textbf{Used for:} melee combat, athletics, endurance tests, physical labor.
  \item \textbf{Typical applications:} lifting, running, climbing, fighting, resisting physical harm.
  \item \textbf{Associated skills:} Athletics, Brawl, Melee, Endurance (see \ref{sec:skills}).
\end{itemize}
\textbf{Rating guide:}
\begin{itemize}
  \item 1: Average person; some physical activity.
  \item 2: Fit; trains or works physically on a regular basis.
  \item 3: Athlete or soldier; excellent condition.
  \item 4: Exceptional athlete; near peak human.
  \item 5: Peak human capability; legendary strength and stamina.
\end{itemize}

\subsection{Wits}
\label{sec:attribute-wits}
\index{Wits}
Mental acuity, perception, quick thinking, and problem-solving.
\begin{itemize}
  \item \textbf{Used for:} investigation, perception, tactics, snap decisions.
  \item \textbf{Typical applications:} spotting details, solving puzzles, planning, reacting quickly.
  \item \textbf{Associated skills:} Investigation, Tactics, Lore, Survival (see \ref{sec:skills}).
\end{itemize}
\textbf{Rating guide:}
\begin{itemize}
  \item 1: Average awareness; sometimes misses important cues.
  \item 2: Observant; notices most relevant details.
  \item 3: Sharp-minded; quick to spot patterns.
  \item 4: Exceptionally perceptive; rarely surprised.
  \item 5: Almost uncanny awareness; sees connections others miss.
\end{itemize}

\subsection{Spirit}
\label{sec:attribute-spirit}
\index{Spirit}
Willpower, intuition, mental resilience, and connection to intangible forces.
\begin{itemize}
  \item \textbf{Used for:} resisting mental effects, intuition, magical aptitude, persistence.
  \item \textbf{Typical applications:} resisting fear, sensing danger, channeling magic, enduring hardship.
  \item \textbf{Associated skills:} Arcana, Ritual (for magic users), Endurance, Insight.
\end{itemize}
\textbf{Rating guide:}
\begin{itemize}
  \item 1: Average willpower; somewhat suggestible.
  \item 2: Strong-minded; resists ordinary pressure.
  \item 3: Very determined; hard to intimidate.
  \item 4: Exceptional will; can inspire others.
  \item 5: Iron will; nearly unshakeable resolve.
\end{itemize}

\subsection{Presence}
\label{sec:attribute-presence}
\index{Presence}
Charisma, social influence, appearance, and force of personality.
\begin{itemize}
  \item \textbf{Used for:} social interactions, leadership, persuasion, intimidation.
  \item \textbf{Typical applications:} negotiating, leading, charming, commanding attention.
  \item \textbf{Associated skills:} Sway, Command, Performance, Deception (see \ref{sec:skills}).
\end{itemize}
\textbf{Rating guide:}
\begin{itemize}
  \item 1: Average presence; does not particularly stand out.
  \item 2: Noticeable; leaves a mild impression.
  \item 3: Charismatic; naturally influential.
  \item 4: Commanding; people listen when you speak.
  \item 5: Magnetic; can sway crowds and redefine the mood of a room.
\end{itemize}

% ============================================================
% SKILL SYSTEM (REVISED)
% ============================================================
\newpage
\section{Skill System}
\label{sec:skills}
\index{Skills}

Skills represent specialized training and expertise. In play, you combine an
Attribute with a relevant Skill to form your dice pool (see \ref{ch:core-mechanics}).

\subsection{Skill Ratings}
\label{sec:skill-ratings}
\index{Skill levels}
\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Rating} & \textbf{Description} \\
\midrule
0 & Untrained --- no formal experience \\
1 & Novice --- basic understanding \\
2 & Competent --- reliable and capable \\
3 & Professional --- expert in the field \\
4 & Master --- renowned specialist \\
5 & Grand Master --- legendary skill \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Skill Categories}
\label{sec:skill-categories}
\index{Combat skills}\index{Physical skills}\index{Social skills}\index{Knowledge skills}\index{Magic skills}
Below are the eighteen core skills, grouped by theme. Your table may rename or reskin
these, but the roles remain similar.

\paragraph{Combat Skills}
\begin{itemize}
  \item \textbf{Melee}: swords, axes, close-quarters weapons.
  \item \textbf{Ranged}: bows, crossbows, thrown weapons, firearms (if present).
  \item \textbf{Brawl}: unarmed combat, grappling, improvised fighting.
  \item \textbf{Tactics}: battlefield strategy, unit coordination, ambush planning.
\end{itemize}

\paragraph{Physical Skills}
\begin{itemize}
  \item \textbf{Athletics}: running, climbing, jumping, swimming.
  \item \textbf{Stealth}: moving unseen, hiding, blending into shadows.
  \item \textbf{Endurance}: resisting fatigue, harsh weather, poison, disease.
  \item \textbf{Craft}: building, repairing, creating tools, art, or structures.
  \item \textbf{Survival}: tracking, foraging, predicting weather, navigating wilderness, animal handling (Nature folded into this skill).
\end{itemize}

\paragraph{Social Skills}
\begin{itemize}
  \item \textbf{Sway}: persuasion, negotiation, charm.
  \item \textbf{Command}: leadership, intimidation, issuing orders.
  \item \textbf{Deception}: lying, bluffing, misdirection.
  \item \textbf{Performance}: entertainment, oration, acting, musical display (used by Cantors for Songs).
  \item \textbf{Insight}: reading emotions, detecting lies, understanding motives (covers social observation and ``reading'' people).
\end{itemize}

\paragraph{Knowledge Skills}
\begin{itemize}
  \item \textbf{Lore}: history, culture, religion, customs, general knowledge (cities, civilizations, scholarly topics).
  \item \textbf{Investigation}: research, deduction, analysis of clues (covers finding and interpreting evidence).
  \item \textbf{Medicine}: healing, anatomy, first aid, treatment.
  % Nature is folded into Survival; no separate Nature skill.
\end{itemize}

\paragraph{Magic Skills}
\begin{itemize}
  \item \textbf{Arcana}: magic theory, rituals (identification), mystical phenomena, free casting (TAGS system), counter-spelling.
  \item \textbf{Ritual}: performing rites, invocations, exorcisms, consecrations -- used by Runekeepers, Invokers, Cantors, and Summoners for active magical workings.
\end{itemize}

\noindent\textbf{Optional Module -- Psionics:} If your table uses the \textit{Psionics} expansion (see \textit{The Gray Wanderer's Grimoire}), a separate \textbf{Psionics} skill is added. It is not part of the core skill list and is used for telepathy, telekinesis, clairvoyance, etc.

% ============================================================
% BUILDING DICE POOLS (unchanged except skill names)
% ============================================================
\newpage
\section{Building Dice Pools}
\label{sec:dice-pools}
\index{Dice pools}

Your dice pool for any action is:
\[
\text{Dice Pool} = \text{Attribute} + \text{Skill}.
\]

You roll that many d10s and interpret the results according to the core rules
(see \ref{ch:core-mechanics}).

\subsection{Choosing the Right Combination}
\label{sec:attribute-skill-pairs}
\index{Attribute!combinations}\index{Skill!combinations}

The same fictional action can often be approached with different Attribute + Skill
pairs, depending on how your character does it. The GM has final say, but the
default is to use the most obvious pairing (e.g., \textit{Body + Melee} for a
straightforward sword swing). Creative alternatives are encouraged when they
fit the fiction.

\begin{itemize}
  \item \textbf{Climbing a wall:}
    \begin{itemize}
      \item \textit{Body + Athletics}: sheer strength and stamina.
      \item \textit{Wits + Athletics}: clever route‑finding and leverage.
      \item \textit{Spirit + Athletics}: grim determination against fear or pain.
    \end{itemize}
  \item \textbf{Persuading a guard:}
    \begin{itemize}
      \item \textit{Presence + Sway}: charm and personality.
      \item \textit{Wits + Sway}: logical, well‑constructed arguments.
      \item \textit{Spirit + Sway}: conviction that burns through doubt.
    \end{itemize}
  \item \textbf{Investigating a crime scene:}
    \begin{itemize}
      \item \textit{Wits + Investigation}: methodical observation and deduction.
      \item \textit{Spirit + Investigation}: intuitive leaps and gut feelings.
      \item \textit{Presence + Investigation}: getting witnesses to open up.
    \end{itemize}
\end{itemize}

\subsection{Creative Combinations}
\label{sec:creative-dice-pools}
With your GM's agreement, you can justify unusual combinations that fit the
fiction. The table below offers common alternatives, but the GM may allow
others based on the scene.

\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Intended Action} & \textbf{Standard Pair} & \textbf{Alternative Pairs (Example)} \\
\midrule
Recall a combat drill & Wits + Lore & Body + Lore (muscle memory) \\
Soothe a patient & Presence + Medicine & Spirit + Medicine (laying on hands) \\
Forge a masterwork & Wits + Craft & Spirit + Craft (inspired artistry) \\
Read weather signs & Wits + Survival & Spirit + Survival (omen‑sense) \\
Feint in a duel & Wits + Melee & Presence + Melee (intimidation feint) \\
Track a spirit & Wits + Survival & Spirit + Lore (follow its legend) \\
Bribe a guard & Presence + Sway & Wits + Streetwise (know the right price) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{GM Guidance for Creative Pairs}
\label{sec:gm-creative-pairs}
\begin{itemize}
  \item \textbf{Ask ``How?''} -- The player must describe \textit{how} they use
        that Attribute and Skill together.
  \item \textbf{Maintain Position} -- A creative pair does not automatically
        improve Position or reduce DV. It just changes the dice pool.
  \item \textbf{Consistency} -- If you allow \textit{Spirit + Craft} for
        visionary work, allow it again in similar circumstances.
  \item \textbf{No Double Dipping} -- The player cannot swap both Attribute
        \textit{and} Skill to max out a single stat; the combination must
        still respect the fiction.
\end{itemize}

\subsection{When to Roll (and When Not To)}
\label{sec:when-to-roll}
Not every action requires a roll. The GM should call for a roll when:
\begin{enumerate}
  \item The outcome is uncertain.
  \item Failure would meaningfully change the situation.
  \item The action is significant (not routine).
\end{enumerate}
For routine actions with no real risk, the group can simply agree that you
succeed. Example: climbing a knotted rope in calm weather -- no roll.
Climbing the same rope while goblins shoot arrows -- roll.

\subsection{Assistance and Teamwork}
\label{sec:assistance-teamwork}
When multiple characters contribute to the same action:
\begin{itemize}
  \item \textbf{Assistance:} One character is primary; up to three others add
        +1 die each by describing how they help (e.g., ``I hold the ladder steady'').
  \item \textbf{Cooperation:} Several characters each roll; the group succeeds
        or fails based on combined results (e.g., all must sneak past a patrol).
  \item \textbf{Complementary:} Different skills cover different parts of a
        larger task (e.g., one distracts a guard, another picks the lock).
\end{itemize}

\emph{Example:} A ranger scales an ice wall using \textbf{Wits + Athletics}
to find the safest holds, then switches to \textbf{Body + Athletics} to muscle over the
lip. The fiction guides which combination fits each moment.

\subsection{Selling the Action -- Intricate Description}
\label{sec:intricate-description}
\index{Intricate description}\index{Selling it}

You are always welcome to state your action plainly. But when you take a moment
to \textbf{sell it} -- to invest in the fiction, reveal what your character cares
about, or show why this moment matters -- the GM may reward you with a mechanical
edge. This is about \emph{roleplay investment}, not word count.

\begin{itemize}
  \item \textbf{Basic description:} ``I attack the goblin with my sword.''
        No bonus.

  \item \textbf{Tactical description:} Use the environment, positioning, or a
        clever tactic. ``I feint high, then drive my shoulder into the goblin's
        shield, shoving it aside to open its flank.''
        The GM may or may not upgrade your Position by one step (e.g., \textit{Controlled}
        $\rightarrow$ \textit{Dominant}) for this roll.

  \item \textbf{Character‑driven description:} Reveal a stake, a bond, a fear,
        or a hard choice. \emph{One heartfelt sentence counts more than a
        paragraph of adjectives.}
        \begin{itemize}
          \item Example: ``I think of my mother, who died because I wasn't brave
                enough to fight. I won't let that happen again. I plant my feet
                and dare the goblin to come through me.''
        \end{itemize}
        The GM may grant a larger benefit -- e.g., upgrade Position \emph{and}
        bank a \textbf{Boon} you can use later, or upgrade Effect by one step.
\end{itemize}

\subsubsection*{What Counts (and What Doesn't)}
\textbf{Counts:}
\begin{itemize}
  \item Stating a clear stake: ``If I fail this lockpick, my sister doesn't eat.''
  \item Invoking a bond mid‑action: ``I owe Ethan my life. I'm not letting him
        face this alone.''
  \item Revealing a fear or hope tied to the roll: ``I'm terrified of fire, but
        I grab the burning beam -- because the children are behind me.''
  \item Making a hard choice visible: ``I could run, but I don't. Running is what
        got me exiled.''
\end{itemize}
\textbf{Does not count:}
\begin{itemize}
  \item Piling on adjectives (``gleaming,'' ``mighty,'' ``ancient'') without stakes.
  \item Passive sensory laundry lists (``the cold wind, the wet stones, the dim light'').
  \item Repeating the same florid description for every roll.
\end{itemize}

\subsubsection*{Guidelines for the GM}
\begin{itemize}
  \item \textbf{No penalty for plain speech.} Never punish a player who is tired,
        shy, or simply prefers quick resolution. The bonus is a gift, not a tax.
  \item \textbf{Be consistent.} Reward genuine investment, not vocabulary. A
        quiet player who delivers one heartfelt sentence deserves the bonus as
        much as an eloquent storyteller.
  \item \textbf{The bonus should fit the fiction.} A desperate, reckless action
        described with real stakes might upgrade Effect but worsen Position --
        let the mechanics mirror the emotion.
  \item \textbf{Invite quiet players.} Ask leading questions: ``What's at stake
        for your character right now?'' or ``What are you afraid will happen if
        you fail?'' Gentle prompts help shy players find their moment.
  \item \textbf{Group ``selling.''} If the whole table contributes to describing a
        group action (a coordinated ambush, a joint ritual), grant the bonus to
        the primary roller or give each participant a small benefit.
\end{itemize}

\subsubsection*{Example}
\textit{Basic:} ``I sneak past the sentry.'' (No bonus, roll as normal.)

\textit{Selling it (tactical):} ``I wait for the wind to rattle the shutters, then
pad across the wet cobblestones -- heel‑toe, heel‑toe -- keeping the stack of
barrels between me and the sentry's gaze.'' The GM nods: ``Great. Because you're
using the wind and the barrels, roll at \textit{Dominant} Position.''

\textit{Selling it (character‑driven):} ``My little brother is in that camp.
I've hidden from every fight since we left home -- but not this time. I creep
forward, not quiet anymore, because he needs to know someone came for him.''
The GM: ``That's the heart of it. Roll at \textit{Dominant} Position, and I'll
bank a Story Beat to make his reaction matter.''

\noindent
\textit{Remember:} The dice still decide success or failure. But a well‑told
action makes the story richer -- and earns you an edge when the dice fall.

% ============================================================
% EARNING XP (unchanged)
% ============================================================
\newpage
\section{Earning Experience Points}
\label{sec:earning-xp}
\index{Experience Points (XP)!earning}

XP represents learning through action. You gain it by engaging meaningfully
with the world: overcoming challenges, taking risks, making hard choices, and
leaning into your character's story.

\subsection{Session Breakdown}
\label{sec:session-xp}
At the end of each session, XP is awarded based on what happened at the table.
A typical breakdown might look like:

\begin{itemize}
  \item \textbf{Base Participation}: +2~XP for attending and contributing.
  \item \textbf{Major Objectives}: +2--4~XP for completing significant story goals.
  \item \textbf{Discoveries}: +1--2~XP for uncovering important lore, locations, or secrets.
  \item \textbf{Hard Choices}: +1--2~XP for making difficult moral or strategic decisions.
  \item \textbf{Story Engagement}: +1--3~XP for embracing complications and twists.
  \item \textbf{Personal Goals}: +1--2~XP for pursuing your character's individual storylines.
\end{itemize}

\emph{Example:} At the end of a session, the party rogue gains:
\begin{itemize}
  \item +2 XP for participation,
  \item +2 XP for helping retrieve a cursed artifact,
  \item +1 XP for pushing a personal rivalry subplot.
\end{itemize}
Total: 5 XP.

\subsection{Game Pace Options}
\label{sec:game-pace}
\index{Game pace}
You and your GM can decide how quickly characters advance, matching the tone
you want:

\begin{center}
\begin{tabular}{cll}
\toprule
\textbf{Mode} & \textbf{XP / Session} & \textbf{Tone} \\
\midrule
Gritty   & 4--6  & Hard choices, slow growth \\
Standard & 6--10 & Balanced progression \\
Epic     & 10--14 & Heroic, rapid development \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Player Tip:} If you want a sweeping, mythic tale, ask for an
\emph{Epic} pace. If you want advancement to feel hard‑won, lean toward
\emph{Gritty}.

\subsection{Arc Completion Bonus}
\label{sec:arc-bonus}
When a major story arc finishes (typically 3--6 sessions), everyone gains an
additional +8--12 XP. One player may receive +2 XP more for a particularly
memorable contribution. This rewards the story as a whole, not just individual
rolls.

\begin{tcolorbox}[colback=yellow!5!white, colframe=yellow!75!black, title=Boons vs. XP]
\index{Boons!conversion}
You may convert 2 Boons $\rightarrow$ 1 XP (once per session, max 2 XP per session via conversion). \textbf{But:} Spending Boons during play is almost always more fun. A re‑rolled failure, an improved Position, or an activated Asset creates a memorable moment. Hoarding Boons for XP conversion starves the table of drama. Spend freely; the XP will come from your adventures.
\end{tcolorbox}

% ============================================================
% SPENDING XP (unchanged except skill names)
% ============================================================
\section{Spending Experience Points}
\label{sec:spending-xp}
\index{Experience Points (XP)!spending}

XP is your currency for growth. You can invest it in three broad areas:

\begin{enumerate}
  \item \textbf{Personal Improvement} (Attributes and Skills)
  \item \textbf{Special Abilities} (Talents, magic, unique tricks)
  \item \textbf{Resources and Influence} (Assets, followers, organizations)
\end{enumerate}

Think of these as three pillars: \textit{what you can do}, \textit{what makes you unique},
and \textit{what you own or command}.

\subsection{Personal Improvement}

\subsubsection*{Attributes}
\index{Attributes!increasing}
Attributes are expensive but powerful. The cost to raise an Attribute is:
\[
\text{XP Cost} = 3 \times \text{new rating}.
\]
Examples:
\begin{itemize}
  \item Raising Body from 2 to 3 costs $3 \times 3 = 9$ XP.
  \item Raising Spirit from 4 to 5 costs $3 \times 5 = 15$ XP.
\end{itemize}
Attribute increases usually require downtime equal to the new rating in days,
spent training, reflecting, or transforming yourself (see \ref{sec:training-time}).

\subsubsection*{Skills}
\index{Skills!increasing}
Skills are cheaper and represent focused practice. The cost is:
\[
\text{XP Cost} = 2 \times \text{new level}.
\]
Examples:
\begin{itemize}
  \item Lore 1 $\rightarrow$ 2 costs $2 \times 2 = 4$ XP.
  \item Melee 3 $\rightarrow$ 4 costs $2 \times 4 = 8$ XP.
\end{itemize}
Skill increases typically require downtime equal to the new level in days.

\medskip
\noindent\textit{Example:} Kara wants to improve her swordsmanship (Melee) from 2 to 3.  
She saves 6 XP and spends three in‑game days training with her mentor. This
training becomes part of the story, not just a line on a sheet.

\subsection{Special Abilities}
\index{Talents}\index{Special abilities}
Special abilities are unique moves, talents, or tricks that define your style.

\begin{description}
  \item[General Abilities (4--8 XP)]  
    Universal benefits, like improved recovery, bonus dice in a common
    situation, or a signature combat technique.  
    \emph{Example:} ``Quick Recovery'' --- heal 1 extra Harm when you rest.

  \item[Cultural Abilities (variable)]  
    Abilities tied to heritage or background. They may require specific
    fictional positioning (see \ref{sec:small-folk-detailed}).  
    \emph{Example:} ``Stone Sense'' --- an instinctive feel for stonework and tunnels.

  \item[Advanced Abilities (12+ XP)]  
    High‑impact features available at higher tiers, often with strong narrative
    hooks or prerequisites (see \ref{sec:tiers}).  
    \emph{Example:} ``Master Diplomat'' --- once per session, you can reroll or
    soften the outcome of a failed major negotiation.
\end{description}

\emph{Example:} A veteran bard purchases \emph{Silver Tongue} for 6 XP,
gaining the ability to sway a hostile crowd once per session. This quickly
becomes their trademark in tense public scenes.

\subsection{Resources and Influence}
\index{Assets}\index{Followers}
See \S\ref{ch:assets-followers} for full rules. As a quick reference:

\begin{description}
  \item[Minor Resource (4 XP, ~1 week)]  
    A small shop, basic workshop, discreet safe house, or minor contact
    network. Offers modest but reliable benefits.

  \item[Standard Resource (8 XP, ~2 weeks)]  
    A decent‑sized business, crew, or operation with real impact. Comes with
    some upkeep and obligations.

  \item[Major Resource (12 XP, ~1 month)]  
    A large enterprise, elite team, or rare capability. Very potent but also
    demanding; often becomes a major part of your character's story.
\end{description}

\noindent\textbf{Player Tip:} Resources pull the camera out from your character
to your \emph{sphere of influence}. A spy network means intrigue. A workshop
means invention. A guild hall means politics.

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=When to Invest in Assets and Followers]
\index{Assets!investment advice}\index{Followers!investment advice}
\centering
\begin{tabularx}{0.95\textwidth}{l l X}
\toprule
\textbf{Player Archetype} & \textbf{Typical Investment} & \textbf{Best For} \\
\midrule
Solo & 0--10\% XP & Reliable personal power, minimal upkeep. \\
Mixed (Balanced) & 15--25\% XP & One solid Follower or Asset (e.g., a scout or a safehouse). \\
Mastermind & 35--55\% XP & Networks, informants, multiple Assets; build a faction. \\
\bottomrule
\end{tabularx}
\vspace{0.2cm}

\textbf{Planning for succession?} A loyal Follower (Cap 3+) can become your successor -- see the \textbf{Legacy Engine} (\S\ref{ch:legacy-engine}). Investing in a potential heir early pays off when your character retires or falls.
\end{tcolorbox}

% ============================================================
% DEVELOPMENT PATHS (unchanged)
% ============================================================
\newpage
\section{Character Development Paths}
\label{sec:development-paths}
\index{Development paths}

How you spend XP shapes your character's growth. These are broad patterns, not
rigid classes.

\begin{center}
\begin{tabular}{p{3cm}p{2.8cm}p{6cm}}
\toprule
\textbf{Path} & \textbf{Typical XP Split} & \textbf{Strengths / Weaknesses} \\
\midrule
Specialist & 70--90\% personal, 0--10\% resources, 0--20\% abilities & Extremely effective in a narrow niche. Vulnerable outside it. \\
Leader     & 50--65\% personal, 15--25\% resources, 15--25\% abilities & Well‑rounded, boosts the group. Rarely the best at one thing. \\
Mastermind & 25--40\% personal, 35--55\% resources, 20--40\% abilities & Solves problems through assets. Personally vulnerable, many moving parts. \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textit{Example (Leader):} A merchant‑prince spends 60\% of XP on personal
skills and social abilities, 20\% on a network of trade contacts, and 20\% on
talents that improve negotiation and command.

These patterns are \emph{guides}, not rules. You can shift between them as your
story evolves.

% ============================================================
% TRAINING & TIME (unchanged)
% ============================================================
\section{Training and Development Time}
\label{sec:training-time}
\index{Training}\index{Downtime}

Most improvements take time in the fiction. Training montages, study scenes,
and building projects are all opportunities for character moments.

\subsection{Standard Time Requirements}
\label{sec:standard-training}
\begin{itemize}
  \item \textbf{Attribute increase:} new rating in days.
  \item \textbf{Skill improvement:} new level in days.
  \item \textbf{Resource acquisition:} about 1 week to 1 month depending on scope.
  \item \textbf{Ability learning:} typically 3--10 days.
\end{itemize}

\subsection{Accelerated Development (Rush Rule)}
\label{sec:rush-rule}
\index{Rush rule}
You \emph{can} attempt to learn faster, but rushing has a price.
\begin{itemize}
  \item The group tracks a short \textbf{Risk Timer } (4 segments) for your crash training.
  \item If it fills, the new capability has flaws:
    \begin{itemize}
      \item Attribute/Skill: temporary --1 die penalty until you retrain properly.
      \item Resource: loyalty issues, partial effectiveness, or hidden problems.
      \item Ability: unreliable effects or quirky side consequences.
    \end{itemize}
\end{itemize}
\noindent\textit{Example:} A wizard crams advanced spellwork into three frantic days.
She gains the ability, but the Risk Timer fills: her spells spark and flicker
until she takes time to train safely.

% ============================================================
% TIERS (unchanged)
% ============================================================
\newpage
\section{Character Progression Tiers}
\label{sec:tiers}
\index{Tiers}\index{Progression}

As you accumulate XP and capabilities, you advance through tiers that represent your growing reputation and influence.

\subsection{Tier I: Novice (0--40 XP)}
\label{sec:tier-i}
\begin{itemize}
\item Learning the ropes, establishing yourself.
\item Local reputation, modest capabilities.
\item Typical assets: basic equipment, a few contacts.
\item \textbf{Position Manifestation:} The difference between life and death; a shift from Desperate to Controlled means ``I live to fight another day.''
\end{itemize}

\subsection{Tier II: Seasoned (41--90 XP)}
\label{sec:tier-ii}
\begin{itemize}
\item Proven capability, reliable in your specialty.
\item Regional reputation, respected in your field.
\item Typical assets: skilled followers, specialized equipment.
\item \textbf{Position Manifestation:} The difference between victory and defeat; a shift from Controlled to Dominant means ``I've earned the right to choose how this ends.''
\end{itemize}

\subsection{Tier III: Veteran (91--150 XP)}
\label{sec:tier-iii}
\begin{itemize}
\item Master of your craft, significant influence.
\item National reputation, can handle major challenges.
\item Typical assets: multiple operations, elite teams.
\item \textbf{Position Manifestation:} The difference between influence and control; a shift from Controlled to Dominant means ``The city hangs on my decision.''
\end{itemize}

\subsection{Tier IV: Paragon (151--220 XP)}
\label{sec:tier-iv}
\begin{itemize}
\item Exceptional capability, strategic influence.
\item Continental influence, shapes regional outcomes.
\item Typical assets: organizations, unique capabilities.
\item \textbf{Position Manifestation:} The difference between shaping events and defining eras; a shift from Controlled to Dominant means ``I am the axis upon which history turns.''
\end{itemize}

\subsection{Tier V: Mythic (221+ XP)}
\label{sec:tier-v}
\begin{itemize}
\item Legendary status, world‑changing influence.
\item Historical reputation, defines eras.
\item Typical assets: nations, legendary artifacts.
\item \textbf{Position Manifestation:} The difference between ending and remaking the world; a shift from Desperate to Controlled means ``The universe waits on my choice.''
\end{itemize}

\subsection{Position Scales with Narrative Weight}
\label{sec:position-narrative-weight}
The mechanical effect of Position (re‑rolling one die, see \ref{ch:core-mechanics}) never changes --- but what it \textit{means} evolves with your tier:
\begin{itemize}
\item \textbf{Tier I:} Position defines personal survival.
\item \textbf{Tier III:} Position shapes regional outcomes.
\item \textbf{Tier V:} Position reshapes reality.
\end{itemize}
The system does not need to scale --- your story does.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, boxrule=0.5pt, arc=2mm]
\textbf{GM Tip:} Never add mechanical bonuses to Position. Instead, deepen the narrative weight:
\begin{itemize}
\item \textbf{Tier I:} ``Your Position shows whether you live or die.''
\item \textbf{Tier V:} ``Your Position shows how the world bends to your will.''
\end{itemize}
The mechanics stay consistent so the narrative can grow as large as your imagination allows.
\end{tcolorbox}

% ============================================================
% ALLIES & FOLLOWERS (unchanged)
% ============================================================
\newpage
\section{Managing Allies and Followers}
\label{sec:allies-followers}
\index{Allies}\index{Followers}

Allies can extend your reach but also create obligations and risks. For full
rules, see \S\ref{ch:assets-followers}.

\subsection{Acquisition Costs}
\label{sec:follower-cost}
A skilled helper's cost in XP is roughly their \emph{capability rating squared}.
\emph{Example:} A capability 3 scout costs $3^2 = 9$ XP.

\subsection{Upkeep Requirements}
\label{sec:follower-upkeep}
Each downtime period, you either:
\begin{itemize}
  \item spend XP equal to their capability rating, \textbf{or}
  \item dedicate a meaningful scene to maintaining the relationship (time, favors, protection).
\end{itemize}

\subsection{Risk Management}
\label{sec:follower-risk}
\begin{itemize}
  \item When big complications arise, your allies can be targeted instead of you.
  \item Allies can handle one significant task off‑screen per downtime, but this often creates extra story fallout (the GM gains a Story Beat).
\end{itemize}

% ============================================================
% SKILL ADVANCEMENT & SYNERGY (updated)
% ============================================================
\section{Skill Advancement and Synergies}
\label{sec:skill-advancement}
\index{Skills!advancement}\index{Synergy}

\subsection{XP Costs}
\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Improvement} & \textbf{XP Cost} \\
\midrule
0 $\rightarrow$ 1 & 2 XP \\
1 $\rightarrow$ 2 & 4 XP \\
2 $\rightarrow$ 3 & 6 XP \\
3 $\rightarrow$ 4 & 8 XP \\
4 $\rightarrow$ 5 & 10 XP \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Training Time}
\label{sec:skill-training-time}
\begin{itemize}
  \item 0 $\rightarrow$ 1: 1 day of practice.
  \item 1 $\rightarrow$ 2: 3 days of training.
  \item 2 $\rightarrow$ 3: 1 week of intensive study.
  \item 3 $\rightarrow$ 4: 2 weeks with a master or rigorous practice.
  \item 4 $\rightarrow$ 5: 1 month of dedicated focus.
\end{itemize}

\subsection{Attribute Limits}
\label{sec:attribute-limits}
You cannot have a Skill rating higher than the relevant Attribute. To advance
a Skill further, you must first raise its primary Attribute (see \ref{sec:core-attributes}).

\subsection{Skill Synergies}
\label{sec:skill-synergies}
When the fiction supports it, well‑justified combinations of skills can grant
+1 die or better Position/Effect. Examples:
\begin{itemize}
  \item \textbf{Combat:} Tactics + Command (battlefield coordination), Melee + Athletics (acrobatic strikes).
  \item \textbf{Social:} Sway + Lore (persuasion backed by knowledge), Deception + Performance (long cons).
  \item \textbf{Exploration:} Investigation + Survival (tracking), Lore + Insight (cultural observation).
\end{itemize}

% ============================================================
% USING SKILLS IN PLAY (updated)
% ============================================================
\section{Using Skills in Play}
\label{sec:using-skills}
\index{Skills!using in play}

\subsection{When to Roll}
\label{sec:when-to-roll}
You roll when:
\begin{itemize}
  \item the outcome is uncertain,
  \item failure would matter to the story, and
  \item the action is significant enough to deserve the spotlight.
\end{itemize}
For routine actions with no real risk, the group can often simply agree that
you succeed.

\subsection{Difficulty and Skill Level}
\label{sec:difficulty-skill}
As your skill rises, routine tasks become trivial. A rough guideline:

\begin{center}
\begin{tabular}{cll}
\toprule
\textbf{Skill Level} & \textbf{Routine Task} & \textbf{Challenging Task} \\
\midrule
0 & Needs a roll; modest DV & Needs a roll; higher DV \\
1 & Often needs a roll & Challenging but fair \\
2 & Usually automatic & Needs a roll; moderate DV \\
3 & Automatic & Needs a roll; interesting stakes \\
4+ & Automatic & Automatic for simple cases \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Reading this as a player:} by Skill 3, basic tasks in your area of
expertise should rarely require rolls. You are there to be impressive, not to
trip over routine jobs.

\subsection{Group Skill Use}
\label{sec:group-skills}
When multiple characters contribute:
\begin{itemize}
  \item \textbf{Assistance:} One character is primary; up to three others add +1 die each by describing how they help (see \ref{ch:core-mechanics}).
  \item \textbf{Cooperation:} Several characters each roll; the group succeeds or fails based on combined results.
  \item \textbf{Complementary:} Different skills cover different parts of a larger task (e.g. one distracts, one picks a lock).
\end{itemize}

% ============================================================
% SKILL CHALLENGES (unchanged)
% ============================================================
\section{Skill Challenges}
\label{sec:skill-challenges}
\index{Skill challenges}\index{Timers}

Some goals are too big for a single roll.

\subsection{Extended Tests}
\label{sec:extended-tests}
For long, involved tasks:
\begin{itemize}
  \item The group sets a progress timer (4--8 segments).
  \item Each significant success fills one or more segments.
  \item Complications can add segments or create new problems.
\end{itemize}

\subsection{Complex Challenges}
\label{sec:complex-challenges}
For multi‑step or multi‑skill situations:
\begin{itemize}
  \item Different skills tackle different phases of the challenge.
  \item Success in one area creates opportunities; failure creates obstacles.
\end{itemize}
\noindent\textit{Example (Heist):} \textbf{Stealth} to infiltrate, \textbf{Craft} to bypass
locks, \textbf{Investigation} to find the vault, \textbf{Deception} to mislead guards.
Each roll advances a \emph{Heist Timer }; complications add heat, suspicion, or
security changes.

% ============================================================
% STRATEGIC ADVANCEMENT (unchanged)
% ============================================================
\section{Strategic Advancement Considerations}
\label{sec:advancement-strategy}
\index{Advancement!strategies}

\subsection{Early Game (Tiers I--II)}
\label{sec:early-game}
\begin{itemize}
  \item Focus on survival and a clear niche.
  \item Raise a key Attribute to 3 early.
  \item Bring core Skills to 2--3.
  \item Establish one or two modest resources or safe havens.
\end{itemize}

\subsection{Mid Game (Tier III)}
\label{sec:mid-game}
\begin{itemize}
  \item Develop signature moves or combinations.
  \item Expand your resource base.
  \item Consider your role in the group and lean into it.
\end{itemize}

\subsection{Late Game (Tiers IV--V)}
\label{sec:late-game}
\begin{itemize}
  \item Choose advanced abilities that reshape the story around you.
  \item Build organizations, movements, or legacies.
  \item Decide how your character will leave their mark on the world.
\end{itemize}

\begin{tcolorbox}[colback=orange!5!white, colframe=orange!75!black, title=Common Advancement Mistakes]
\noindent\textbf{Avoid these pitfalls:}
\begin{itemize}
  \item \textbf{Over‑specializing too early:} A Melee 5 character with Body 3 is fragile and incapable outside combat.
  \item \textbf{Forgetting training time:} You cannot learn a new skill in the middle of a dungeon without justification.
  \item \textbf{Neglecting resources:} A few well‑chosen assets can solve problems no amount of personal skill can.
  \item \textbf{Hoarding XP:} Spend your XP! Characters who never improve become irrelevant to the escalating threats.
\end{itemize}
\end{tcolorbox}

% ============================================================
% PRACTICAL EXAMPLES (updated skills)
% ============================================================
\section{Practical Examples}
\label{sec:practical-examples}
\index{Examples}

\subsection{Combat Example}
\label{sec:combat-example}
A warrior with Body 4 and Melee 3 attacks:
\begin{itemize}
  \item Dice pool: $4 + 3 = 7$d10.
  \item Opponent's defenses and the fiction help set the difficulty (see \ref{ch:core-mechanics}).
  \item Strong successes translate into higher Effect or better Position.
\end{itemize}

\subsection{Social Example}
\label{sec:social-example}
A diplomat with Presence 3 and Sway 2 negotiates:
\begin{itemize}
  \item Dice pool: $3 + 2 = 5$d10.
  \item Position might be \emph{Controlled} if circumstances favor them, or worse if under suspicion.
  \item Consequences for failure could be mistrust, delays, or new complications.
\end{itemize}

\subsection{Exploration Example}
\label{sec:exploration-example}
A scout with Wits 3 and Survival 2 searches for tracks:
\begin{itemize}
  \item Dice pool: $3 + 2 = 5$d10.
  \item Easier tasks (fresh tracks) are more forgiving; older or obscured tracks are harder.
  \item Partial successes might reveal some information but not all, or come with side effects (e.g., you find the tracks but disturb a nest of predators).
\end{itemize}

\noindent Your Attributes and Skills determine not just \emph{if} you succeed, but
\emph{how} you tackle problems. Lean into combinations that express your
character's style.

% ============================================================
% QUICK REFERENCE BOXES (updated)
% ============================================================
\begin{tcolorbox}[colback=green!5!white,colframe=green!75!black,title=XP Planning Guide,fonttitle=\bfseries]
\textbf{Early Tier Priorities:}
\begin{itemize}
  \item Raise a core Attribute to 3 (9 XP).
  \item Bring key Skills to 2--3 (4--6 XP each).
  \item Claim 1--2 minor resources (4 XP each).
\end{itemize}

\textbf{Mid Tier Expansion:}
\begin{itemize}
  \item Push core Attributes toward 4 (12 XP).
  \item Specialize key Skills to 4 (8 XP).
  \item Add standard resources (8 XP each).
  \item Explore cultural or thematic abilities (6--10 XP).
\end{itemize}

\textbf{Late Tier Mastery:}
\begin{itemize}
  \item Purchase capstone abilities (12+ XP).
  \item Establish major resources (12 XP).
  \item Launch legacy projects, organizations, or movements.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Attributes and Skills Quick Reference,fonttitle=\bfseries]
\textbf{Attributes (1--5):}
\begin{itemize}
  \item \textbf{Body}: physical capability
  \item \textbf{Wits}: mental acuity
  \item \textbf{Spirit}: willpower and inner strength
  \item \textbf{Presence}: social influence
\end{itemize}

\textbf{Skill Levels:}
\begin{itemize}
  \item 0: Untrained \quad 1: Novice \quad 2: Competent
  \item 3: Professional \quad 4: Master \quad 5: Grand Master
\end{itemize}

\textbf{Core Skills (18):}
\begin{itemize}
  \item Combat: Melee, Ranged, Brawl, Tactics
  \item Physical: Athletics, Stealth, Endurance, Craft, Survival
  \item Social: Sway, Command, Deception, Performance, Insight
  \item Knowledge: Lore, Investigation, Medicine
  \item Magic: Arcana, Ritual
\end{itemize}

\textbf{Dice Pool:} Attribute $+$ Skill d10s.

\textbf{Skill Improvement:} New level $\times$ 2 XP (plus training time).

\textbf{Specialization:} At Skill 3+, you are a known expert in that field.

\textbf{Synergy:} Well‑justified combinations of skills can grant +1 die or
better Position/Effect when the fiction supports it.
\end{tcolorbox}