\chapter{Appendices}

% Let LaTeX handle the lettering automatically.
% Each section below becomes Appendix A, B, C, etc.

\section{Player Tools and Quick Reference}
\label{app:player-tools}

This appendix provides lightweight, table‑friendly tools for tracking character progress, campaign state, and moment‑to‑moment rules. Print, photocopy, or copy to index cards for use at the table.

\subsection{Physical Tracker Recommendations}

Many groups find that physical, tactile trackers speed up play and reduce mental load. Consider:

\begin{itemize}
  \item \textbf{Poker chips or glass beads} for Boons and Story Beats (SB).  
  \emph{Keep a small bowl of chips in the center of the table -- spend and earn visibly.}
  \item \textbf{Dry‑erase boards or laminated sheets} for timers (4/6/8 segment circles).  
  \emph{Use a wet‑erase marker; draw a new timer each scene.}
  \item \textbf{Index cards (3×5)} for Assets and Followers.  
  \emph{Write the Asset's name, free use, and surge on one side; flip when used.}
  \item \textbf{Colored paperclips or binder clips} on character sheets for Fatigue and Harm tracks.  
  \emph{Slide the clip up or down -- no erasing.}
  \item \textbf{Printed ``timer wheels''} (a circle divided into 4/6/8 wedges) placed inside a page protector.  
  \emph{Mark wedges with a dry‑erase marker, erase with tissue.}
  \item \textbf{D10 dice} to track temporary modifiers (Position, Effect, or Momentum).  
  \emph{Place the die next to the relevant tracker and tick up or down.}
\end{itemize}

\subsection{Resolution Index Card}

\begin{tcolorbox}[colback=white,colframe=black,left=1em,right=1em,top=0.5em,bottom=0.5em,breakable]
\textbf{Declare Action → Build Dice Pool (Attribute + Skill) → Roll → Count Successes}

\medskip
0 Successes = \textbf{Miss} (GM gains SB, consequences) \\
1--3 Successes = \textbf{Partial} (success + complication) \\
4+ Successes = \textbf{Full Success} (clear win, no complication)

\smallskip
\textbf{Position} (Risk) + \textbf{Effect} (Impact) may shift outcomes.

\smallskip
\textbf{SB (Story Beats)} fuel abilities, magic, momentum, and narrative pushes.
\end{tcolorbox}

\subsection{Fatigue and Harm Tracker}

\begin{tcolorbox}[colback=white,colframe=black,left=1em,right=1em,top=0.5em,bottom=0.5em,breakable]
\textbf{Fatigue:} \_\_\_\_\_\_\_\_  (Max = Body) \\
\textbf{Mental Strain:} \_\_\_\_\_\_\_\_ \\
\textbf{Harm:} [~] Harm 1 (--1 die) [~] Harm 2 (--2 dice) [~] Harm 3 (Incapacitated)

\medskip
When tracks fill:
\begin{itemize}
    \item Fatigue overflow → increase Harm by 1, clear Fatigue.
    \item Mental Strain → causes instability or disadvantage.
    \item Harm 3 → Last Stand or Patron's Claim (see ``Optional Player‑Character Death'' in \S\ref{sec:pc-death-optional}).
\end{itemize}
\end{tcolorbox}

\subsection{SB and Momentum Tracking}

\begin{tcolorbox}[colback=white,colframe=black,left=1em,right=1em,top=0.5em,bottom=0.5em,breakable]
\textbf{Current SB (GM pool):} \_\_\_\_\_\_ \\
\textbf{Momentum (party pool):} \_\_\_\_\_\_  (max 2) \\
\textbf{Used this session:} \_\_\_\_\_\_
\end{tcolorbox}

\subsection{Talent Progression}

\centering
\begin{tabular}{|c|c|}
\hline
\textbf{XP Cost} & \textbf{Talent Type / Impact} \\
\hline
2 XP & Minor Talent / Perk upgrade / situational bonus \\
4 XP & Major Talent / specialization / strong permanent effect \\
6 XP & Access talent (Spellcraft, Familiar, Codex, Patron's Symbol) \\
8 XP & Prestige Talent / campaign‑defining ability \\
\hline
\end{tabular}
\par

\subsection{Character Goals Card}

\begin{tcolorbox}[colback=white,colframe=black,width=\textwidth,left=1em,right=1em,breakable]
\textbf{Name:} \_\_\_\_\_\_\_\_\_\_\_\_ \\
\textbf{Core Drive:} \_\_\_\_\_\_\_\_\_\_\_\_ \\
\textbf{Short-Term Goal (1--2 sessions):} \_\_\_\_\_\_\_\_\_\_\_\_ \\
\textbf{Long-Term Goal (arc):} \_\_\_\_\_\_\_\_\_\_\_\_ \\
\textbf{Background Complication:} \_\_\_\_\_\_\_\_\_\_\_\_ \\
\textbf{Relationship That Matters:} \_\_\_\_\_\_\_\_\_\_\_\_
\end{tcolorbox}

\emph{Physical tip:} Print this card on cardstock, fold in half, and stand it in front of you as a reminder.

\subsection{Faction Attitude Card}

\begin{tcolorbox}[colback=white,colframe=black,left=1em,right=1em,top=0.5em,bottom=0.5em,breakable]
\textbf{Faction Name:} \_\_\_\_\_\_\_\_\_ \\
Attitude: --3 --2 --1 0 +1 +2 +3 \\
Notes: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\end{tcolorbox}

\subsection{Player Orientation Card (Quick Principles)}

\begin{tcolorbox}[colback=white,colframe=black,left=1em,right=1em,breakable]
\textbf{1. Declare Bold Action} --- Drive fiction forward. \\
\textbf{2. Accept Consequences} --- Complications create story. \\
\textbf{3. Spend Story Beats} --- Make moments matter. \\
\textbf{4. Use Boons} --- Re‑roll, improve Position, activate Assets. \\
\textbf{5. Share the Spotlight} --- Assist, bond, set up allies.
\end{tcolorbox}

\subsection{Class / Archetype Resource Quick Reference}

\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Archetype} & \textbf{Core Resource} \\
\hline
Fighter / Warrior & Maneuvers, Fatigue, Weapon Mastery \\
Rogue / Scout & Positioning, Stealth, Backstab \\
Caster (Freeform) & Spellcraft, Elemental Tags, Backlash \\
Runekeeper & Rites, Obligation, Codex \\
Invoker & Rituals, Patron's Symbol, Crack the Seal \\
Cantor & Songs, Corruption Timer , Repertoire \\
Summoner & Leash Timer , Spirit Bond, Pact \\
\hline
\end{tabular}
\par

\subsection{How to Engage with the GM (Player Prompts)}

\begin{itemize}
    \item \textbf{Ask:} ``What do I need to roll?'' -- get the DV and Position.
    \item \textbf{Propose:} ``Can I spend SB to achieve X?'' -- negotiate narrative pushes.
    \item \textbf{Declare:} ``Here's my intent and method.'' -- always state what you want to happen.
    \item \textbf{Offer:} ``I'll take a complication if Y happens.'' -- invite Devil's Bargains.
    \item \textbf{Clarify:} ``What's at risk if I fail?'' -- know the stakes before rolling.
\end{itemize}

\subsection{Downtime Project Tracker}

\begin{tcolorbox}[colback=white,colframe=black,left=1em,right=1em,breakable]
\textbf{Name:} \_\_\_\_\_\_\_\_\_\_\_  \qquad \textbf{Segments:} [ ] [ ] [ ] [ ] [ ] [ ] \\
\textbf{Goal:} \_\_\_\_\_\_\_\_\_\_\_ \\
\textbf{Obstacles / Resources Needed:} \_\_\_\_\_\_\_\_
\end{tcolorbox}

\emph{Physical tip:} Use a 6‑segment whiteboard timer. Each downtime, erase one segment and roll for progress.

\subsection{What Consequence Do You Accept? (Menu for GM)}
\label{app:consequence-menu}

When a player fails or partially succeeds, the GM may offer a choice from this list (or invent a custom one):
\begin{itemize}
    \item Harm (level 1 or 2)
    \item Fatigue (1 or 2)
    \item Lost Advantage (Position worsens, a timer advances)
    \item Time Delays (a timer ticks, a deadline tightens)
    \item Exposure / Heat (attention from enemies, bounty)
    \item Resource Depletion (supply timer tick, gear compromised)
    \item Relationship Strain (a bond or faction attitude worsens)
\end{itemize}

\subsection{Ways to Shine in Fate's Edge}

\begin{itemize}
    \item Tie actions to your drives and backstory -- invoke your Background Boon.
    \item Offer complications and accept them -- earn Boons.
    \item Spend SB to escalate or transform the scene -- make failure memorable.
    \item Use your class resource early and often -- don't hoard.
    \item Connect scenes to campaign themes -- name the faction or place.
    \item Make other characters look cool -- assist, set up, pass the spotlight.
\end{itemize}

\subsection{Blank Player Index Cards (for quick notes)}

\begin{center}
\fbox{\parbox{0.75\textwidth}{\vspace{0.75in} \centering Name / Boons / Fatigue / Harm}}
\end{center}

\begin{center}
\fbox{\parbox{0.75\textwidth}{\vspace{0.75in} \centering Timers / Obligation / Scene Notes}}
\end{center}

\subsection{Final Reminder}

\textbf{Your character sheet is not your character.}  
It is a tool to help you express who they have been, who they are becoming, and how their choices change the world.

Use these tools to make that expression vivid, consequential, and memorable.

\clearpage

\section{Magic Systems Reference}
\label{app:magic-ref}

Magic in Fate's Edge is plural.  
Different traditions channel power through different metaphysical structures, but all obey the same truth:

\begin{center}
\emph{Power has a cost, and magic shapes the one who wields it.}
\end{center}

This appendix provides quick‑reference tools for players using magic‑affiliated systems, talismans, rites, and occult practices.

\subsection{Core Magical Principles}
\begin{tcolorbox}[colback=white,colframe=black,left=1em,right=1em,breakable]
\begin{itemize}[noitemsep]
    \item Magic expresses \textbf{identity and belief}.
    \item Casting always creates \textbf{consequences somewhere}.
    \item Magic rewrites reality through \textbf{intention, structure, or sacrifice}.
    \item Excessive power erodes \textbf{choice} (see Iconic Ascension sidebar in the full rules).
\end{itemize}
\end{tcolorbox}

\subsection{Universal Casting Reference}
\begin{tcolorbox}[colback=gray!5,colframe=gray!60,breakable]
\textbf{Cast Declaration Template:}
\begin{enumerate}[noitemsep]
    \item \textbf{Intent} --- What change do you seek?
    \item \textbf{Shape} --- How does your tradition express it?
    \item \textbf{Cost} --- What do you risk or expend?
    \item \textbf{Resolution} --- Test; apply Position/Effect rules.
    \item \textbf{Backwash} --- Mark Fatigue, SB pressure, taboo, instability, or bargain.
\end{enumerate}
\end{tcolorbox}

\subsection{Magic Tradition Quick Reference}

\begin{description}
  \item[Freeform Arcana (Wizardry)] \textit{Creative, rational, deliberate shaping.} Tools: Glyphs, structures, metaphors. Costs: Fatigue, reality recoil, SB instability. \textbf{Play Cue:} Diagram your spell or state its symbolism. Big spells distort fate -- accept SB spikes.
  
  \item[Innate Arcana (Sorcery)] \textit{Emotional, instinctive, bloodline‑linked.} Tools: Breath, impulse, resonance. Costs: Fatigue, unpredictable backlash. \textbf{Play Cue:} Treat magic like a reflex. Push your limits -- expect consequences.
  
  \item[Invocation (Priest / Paladin / Medium)] \textit{Petition, prayer, pact, rite.} Tools: Symbols, vows, places of power. Costs: Taboo pressures, obligation timers. \textbf{Play Cue:} Your magic is not your own -- frame it as a request. Violating your ethos draws consequences.
  
  \item[Runic Consecration (Runekeeping)] \textit{Methods, seals, contracts.} Tools: Ritual steps, codices, binding clauses. Costs: Hunger timers, escalation loops. \textbf{Play Cue:} Magic is \emph{procedure}. Expect the system to push back.
  
  \item[Shaping / Druidic Power] \textit{Resonance with place, element, cycle.} Tools: Marks, offerings, seasonal rites. Costs: Environmental echo, territorial consequence. \textbf{Play Cue:} Declare what the land wants -- then bargain.
  
  \item[Pact Binding (Warlock / Avatar)] \textit{Exchange, hunger, transformation.} Tools: Tokens, signatures, reflection marks. Costs: Autonomy loss, ascension risk. \textbf{Play Cue:} Ask: ``What does my patron get?'' Debt becomes plot.
  
  \item[Summoning (Medium / Necromancer / Shaman)] \textit{Invocation of entities, binding, vassal action.} Tools: Circles, bargains, lanterns, ancestral calls. Costs: Attention timers, breach risk. \textbf{Play Cue:} Give the summoned a personality -- their aid comes with demands.
\end{description}

\subsection{Mutual Rules of Magical Cost}
\begin{center}
\textbf{No magic is free.}
\end{center}

\centering
\small
\begin{tabular}{|c|p{9cm}|}
\hline
\textbf{Cost Type} & \textbf{Examples in Play} \\
\hline
Fatigue & Exhaustion surge, shaking focus, body strain \\
SB Pressure & GM spikes complications or timers (elemental backlash, Patron omens) \\
Taboo / Obligation & Commandments, rites, punishment for misuse, Patron demands \\
Attention / Backwash & Spirits notice; cosmic ledger ticks; faction awareness \\
Anomaly / Distortion & Places react, reality warps, rituals echo -- scene tags change \\
Identity Erosion & Compelled acts, alignment tilt, Iconic Ascension risk \\
\hline
\end{tabular}
\par

\subsection{Cross‑System Synergies}
All magic systems can blend. A Runekeeper may emulate a Cleric or Druid; an Invoker may operate as a Paladin or Prophet; a Sorcerer may look like a Witch, Elementalist, or Psion.

When combining systems:
\begin{enumerate}
    \item Choose your \textbf{primary expression} (belief, intellect, instinct, ritual).
    \item Describe spells through that lens.
    \item Let costs reflect your system.
    \item Track consequences separately -- mixing systems increases risk.
\end{enumerate}

\subsection{GM Reminders for Magic}
\begin{itemize}
    \item Magic creates story -- reward its use with consequences and momentum.
    \item Players should narrate method, not just effect.
    \item Big moments should change scenes, not just numbers.
    \item Magic should tempt the caster -- echo their drives and flaws.
\end{itemize}

\subsection{Player Reminders for Magic}
\begin{itemize}
    \item Lean into cost -- magic feels significant when it hurts.
    \item Make magic personal -- describe what it looks and feels like.
    \item Tie powers to identity, belief, trauma, duty, or legacy.
    \item Let magic change your character over time.
\end{itemize}

\clearpage

\section{Pre‑made Spells with TAGS and DV Math}
\label{app:spell-examples}

The following examples illustrate how TAGS, DV scaling, and spell intent interact within the casting system. Use them as inspiration or as templates for your own spells.

\subsection{How to Read a Spell Entry}
Each entry includes:
\begin{itemize}
  \item \textbf{Name} and \textbf{Tradition} (optional).
  \item \textbf{TAGS} -- the components that define the effect.
  \item \textbf{Effect} -- plain‑language outcome.
  \item \textbf{DV} -- base calculation (1 + number of TAGS; dangerous tags add +2).
  \item \textbf{Notes} -- typical range, duration, or special considerations.
\end{itemize}

\subsection{Tier I: Minor Utility (DV 2--3)}

\paragraph{Simple Light} (Air + Light) \\
\textbf{TAGS:} {[ILLUMINATE]} \\
\textbf{Effect:} Create a small sustained light (torch‑brightness) for the scene. \\
\textbf{DV:} 1 + 1 = \textbf{2}. \\
\textbf{Notes:} No backlash risk; lasts until dismissed.

\paragraph{HEAL Minor Wounds} (Life + HEAL) \\
\textbf{TAGS:} {[HEAL]} \\
\textbf{Effect:} Close a small cut or bruise; reduce 1 Fatigue (Minor). \\
\textbf{DV:} 1 + 1 = \textbf{2}. \\
\textbf{Notes:} Requires touch.

\begin{tcolorbox}[
    colback=parchment,
    colframe=bronze,
    sharp corners,
    breakable,
    enhanced,
    title=Healing Harm and Fatigue,
    fonttitle=\bfseries,
    attach boxed title to top left={yshift=-2mm,xshift=4mm},
    boxed title style={colback=bronze, colframe=bronze, sharp corners, rounded corners=north}
]

\textbf{HEAL as a Cantrip.} The \texttt{HEAL} tag (Life) can be cast as a single‑action cantrip (DV 2). On a success, you remove \textbf{1 Fatigue} from yourself or a touched ally. This cannot heal Harm.

\textbf{Healing Harm (1 level).} To restore a level of Harm, you must include the \texttt{HEAL} tag and either:
\begin{itemize}
    \item add a \textbf{+2 DV modifier} to the spell, or
    \item combine \texttt{HEAL} with a second appropriate tag (e.g., \texttt{Purify}, \texttt{Strengthen}, \texttt{Cleanse}).
\end{itemize}
The GM sets the final DV based on the tags and the wound's severity.

\textit{Why?} Fatigue is exhaustion -- a quick charm can ease tired muscles. Harm is injury -- closing wounds demands more power and risk. This keeps healing meaningful without overshadowing martial endurance.

\end{tcolorbox}

\paragraph{Message on Wind} (Air + Veil) \\
\textbf{TAGS:} {[WHISPER]} \\
\textbf{Effect:} Carry a short whispered message 100 feet on wind currents. \\
\textbf{DV:} 1 + 1 = \textbf{2}. \\
\textbf{Notes:} One‑way communication; 30 seconds duration.

\subsection{Tier II: Combat Utility (DV 3--4)}

\paragraph{Blade Flame} (Fire + Strike) \\
\textbf{TAGS:} {[BURNING][STRIKE]} \\
\textbf{Effect:} Coat a weapon in fire, adding +1 Harm and {[BURN]} to the next melee attack. \\
\textbf{DV:} 1 + 1 + 1 = \textbf{3}. \\
\textbf{Notes:} Lasts one exchange.

\paragraph{Shield of Air} (Air + Wall) \\
\textbf{TAGS:} {[WARD][DEFEND]} \\
\textbf{Effect:} Create a compressed air barrier; gain Position +1 on next defense. \\
\textbf{DV:} 1 + 1 + 1 = \textbf{3}. \\
\textbf{Notes:} One use only.

\paragraph{Fleet Step} (Air + Leap) \\
\textbf{TAGS:} {[HASTE][MOVE]} \\
\textbf{Effect:} Double movement speed for one exchange. \\
\textbf{DV:} 1 + 1 + 1 = \textbf{3}. \\
\textbf{Notes:} Requires open area.

\subsection{Tier III: Area Control (DV 4--5)}

\paragraph{Wall of Stone} (Earth + Wall) \\
\textbf{TAGS:} {[BARRIER][FORTIFY]} \\
\textbf{Effect:} Summon a 10‑foot stone wall (Near; scene duration). \\
\textbf{DV:} 1 + 1 + 1 = \textbf{3}. \emph{Dangerous DV:} \textbf{4} (combat blocking). \\
\textbf{Notes:} Provides cover; obstructs movement.

\paragraph{Sleep's Gentle Touch} (Water + Mind) \\
\textbf{TAGS:} {[DREAM][COMMAND]} \\
\textbf{Effect:} One target falls asleep (scene duration; Resist: Spirit + Resolve DV 2). \\
\textbf{DV:} 1 + 1 + 1 = \textbf{3}. \emph{Dangerous DV:} \textbf{4} (combat use). \\
\textbf{Notes:} Harm wakes target.

\paragraph{Flame Cloak} (Fire + Area) \\
\textbf{TAGS:} {[BURNING][AREA]} \\
\textbf{Effect:} 6‑foot radius of intense heat; enemies suffer --1 die to approach. \\
\textbf{DV:} 1 + 1 + 1 = \textbf{3}. \emph{Dangerous DV:} \textbf{4} (if allies are near). \\
\textbf{Notes:} Creates difficult terrain.

\subsection{Tier IV: Battlefield Shaping (DV 5--6)}

\paragraph{Earth's Tremor} (Earth + Force) \\
\textbf{TAGS:} {[QUAKE][AREA]} \\
\textbf{Effect:} 20‑foot tremor; enemies test Body + Athletics (DV 3) or fall prone. \\
\textbf{DV:} 1 + 1 + 1 = \textbf{3}. \emph{Dangerous DV:} \textbf{5}. \\
\textbf{Notes:} Destabilizes structures; affects all in area.

\paragraph{Storm's Fury} (Air + Storm) \\
\textbf{TAGS:} {[STORM][AREA][FORCE]} \\
\textbf{Effect:} 30‑foot storm cloud; lightning strikes random targets (Harm 2). \\
\textbf{DV:} 1 + 1 + 1 + 1 = \textbf{4}. \emph{Dangerous DV:} \textbf{6}. \\
\textbf{Notes:} Lasts scene; reduces visibility.

\paragraph{Healing Circle} (Life + Area) \\
\textbf{TAGS:} {[HEAL][AREA][CLEANSE]} \\
\textbf{Effect:} Allies in 15‑foot radius heal 1 Harm and clear 1 Fatigue. \\
\textbf{DV:} 1 + 1 + 1 + 1 = \textbf{4}. \\
\textbf{Notes:} Major positive effect warrants high DV.

\subsection{Tier V: Domain‑Level Effects (DV 6--7)}

\paragraph{Gate of Passage} (Air + Fold) \\
\textbf{TAGS:} {[GATE][TRANSPORT][AREA]} \\
\textbf{Effect:} Open a two‑way portal 20 feet wide linking two locations. \\
\textbf{DV:} 1 + 2 + 1 + 1 = \textbf{5}. \emph{Dangerous DV:} \textbf{7}. \\
\textbf{Notes:} Requires concentration; may destabilize.

\paragraph{Army of Stone} (Earth + Summon) \\
\textbf{TAGS:} {[CONJURE][AREA][GUARD]} \\
\textbf{Effect:} Summon 4 stone guardians (Cap 2) to follow simple commands. \\
\textbf{DV:} 1 + 2 + 1 + 1 = \textbf{5}. \emph{Dangerous DV:} \textbf{6}. \\
\textbf{Notes:} Scene duration; concentration.

\paragraph{Mind's Dominion} (Fate + Mind) \\
\textbf{TAGS:} {[COMMAND][AREA][CURSE]} \\
\textbf{Effect:} All enemies in 25‑foot radius test Spirit + Resolve (DV 4) or obey a single command. \\
\textbf{DV:} 1 + 1 + 1 + 2 = \textbf{5}. \emph{Dangerous DV:} \textbf{7}. \\
\textbf{Notes:} Significant ethical risk; backlash likely.

\subsection{Signature / Personal Spells (Custom)}

\paragraph{The Warden's Bond} (Life + Bind) -- Healing Cantor \\
\textbf{TAGS:} {[HEAL][BIND][GUARD]} \\
\textbf{Effect:} Heal 2 Harm and link caster to target -- the next Harm they take, caster suffers 1 instead. \\
\textbf{DV:} 1 + 1 + 1 + 1 = \textbf{4}. \\
\textbf{Notes:} Sacrificial protection.

\paragraph{Shadow's Step} (Air + Veil) -- Trickster's Path \\
\textbf{TAGS:} {[VEIL][MOVE][LEAP]} \\
\textbf{Effect:} Become invisible and teleport 30 feet; gain +2 dice to next Stealth action. \\
\textbf{DV:} 1 + 1 + 1 + 1 = \textbf{4}. \\
\textbf{Notes:} Requires shadow or dim light.

\paragraph{Forge's Truth} (Fire + Reveal) -- Artificer's Craft \\
\textbf{TAGS:} {[REVEAL][HEAL][FORTIFY]} \\
\textbf{Effect:} See flaws in an object and repair as if 2 Skill levels higher (scene). \\
\textbf{DV:} 1 + 1 + 1 + 1 = \textbf{4}. \\
\textbf{Notes:} Works on magical items (with risk).

\subsection{Backlash and Push Guidelines}

\subsubsection{Backlash Examples by Tier}
\begin{itemize}
    \item \textbf{Tier I--II:} Minor fatigue, brief elemental effect (smoke, chill, static).
    \item \textbf{Tier III--IV:} Harm 1, temporary Condition, environmental hazard.
    \item \textbf{Tier V:} Harm 2, major Condition, significant environmental change.
\end{itemize}

\subsubsection{Push Options (Accelerate or Amplify)}
\begin{itemize}
    \item \textbf{+1 Effect:} Increase radius, duration, or potency one step.
    \item \textbf{+1 die:} Improve casting roll (cost: +1 Fatigue).
    \item \textbf{Instant:} Resolve immediately (cost: +1 Harm).
    \item \textbf{Sustain:} Maintain effect with concentration (+1 Fatigue/round).
\end{itemize}

\noindent This framework provides concrete spell examples and demonstrates how TAGS, DV scaling, and ambition interact within the magic system.

\clearpage

\section{Free Casting (TAGS System)}\label{sec:free_casting_tags}
\index{Freeform magic!TAGS}\index{Tags}

Some casters shape raw forces through shared arcane grammar known as \textbf{TAGS}. Construct a spell at the table using a short phrase of TAGS.

\subsection{Spell Structure}
\textbf{Intent} + \textbf{Target} + \textbf{Tags} = effect. Example: \textit{``I unleash Burning • Area • Force against the marauders.''}

\subsection{Base Difficulty Value (DV)}
\[
\text{DV} = 1 + \text{number of TAGS}
\]
Dangerous TAGS (Teleportation, Transformation, Domination) add +2 DV instead of +1.

\subsection{Casting Roll}
Roll \textbf{Wits + Arcana} (or similar). Success = spell works. Failure or any 1 = Backlash.

\subsection{Backlash for Free Casters}\index{Backlash!Free Caster}
Choose one:
\begin{itemize}
    \item Harm 2 (Arcane)
    \item +2 Fatigue
    \item Corruption +1
    \item Catastrophic side effect (GM describes)
\end{itemize}
If the spell includes a ``Dangerous'' TAG, Backlash triggers on \emph{mixed} results as well.

\subsection{TAG Library (Abridged)}\label{subsec:tag_library}
\index{Tags!List}

\centering
\small
\begin{longtable}{p{3.5cm} p{10cm}}
\caption{Common TAGS}\label{tab:tags_abridged}\\
\toprule
\textbf{TAG Category} & \textbf{Examples} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{{\bfseries Continued from previous page}} \\
\toprule
\textbf{TAG Category} & \textbf{Examples} \\
\midrule
\endhead
\midrule
\multicolumn{2}{r}{{Continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot
Elemental & Burning, Freezing, Storm, Stone, Wave, Wind, Earth, Fire, Water, Air, Lightning, Ash \\
Force & Force, Area, Strike, Wall, Bind, Dispel, Push, Pull, Crush, Shield \\
Mind \& Veil & Veil, Scry, Memory, Command, Fear, Glamour, Dream, Whisper, Reveal, Hide \\
Life \& Body & HEAL, Purify, Strengthen, Waken, Beast, Heal, Poison, Rot, Numb, Sicken \\
Space \& Motion (dangerous, +2 DV) & Leap, Fold, Gate, Gravity, Phase, Teleport, Blink, Anchor \\
Creation \& Transformation (dangerous, +2 DV) & Create, Summon, Transmute, Animate, Conjure, Shape, Graft, Fleshcraft \\
Illusion \& Deception & Mirage, Echo, Mask, Copy, Phantasm, Glimmer \\
Protection \& Warding & Ward, Defend, Seal, Barrier, Sanctuary, Negate \\
Time (rare, dangerous, +2 DV) & Haste, Slow, Stop, Rewind, Age \\
Curse \& Hex & Curse, Jinx, Bind, Misfortune, Taint, Corrupt \\
Information & Know, Sense, Locate, Recall, Translate \\
\end{longtable}
\par

\subsection{Example Spells by DV}
\begin{itemize}
    \item \textbf{Light} [Illuminate] -- DV 2 (1 tag)
    \item \textbf{Message on Wind} [Whisper] -- DV 2
    \item \textbf{Shield of Air} [Ward] [Defend] -- DV 3
    \item \textbf{Flame Cloak} [Burning] [Area] -- DV 3
    \item \textbf{Wall of Stone} [Barrier] [Fortify] -- dangerous? No, DV 3 (two tags) or DV 4 (if GM rules combat blocking as dangerous)
    \item \textbf{Teleport} [Leap] [Veil] -- DV 1 + 2 (dangerous) + 1 = 4, but Leap is already dangerous: [Leap] alone is dangerous (+2) so DV 3.
\end{itemize}

%\chapter{Additional Patrons and Complete Tables}
%\label{app:additional-patrons}
%
%The following pages contain the full entries for patrons not included in the main chapter. These powers are still part of the world; GMs may introduce them as needed. The complete 24‑patron gift map and fear influence table are also reproduced here for reference.
%
%% --- Patron Archetypes and Build Themes ---
%\section{Patron Archetypes and Build Themes}
%\label{app:patron-archetypes}
%
%The table below summarizes each patron's mechanical and narrative niche. Use it to quickly find a patron that matches your character concept.
%
%\centering
%\small
%\renewcommand{\arraystretch}{1.2}
%\begin{longtable}{@{}p{4.0cm}p{4.5cm}p{7.0cm}@{}}
%\toprule
%\textbf{Patron} & \textbf{Archetype / Build Theme} & \textbf{Signature Mechanics} \\
%\midrule
%\endfirsthead
%
%\multicolumn{3}{c}{{\bfseries Continued from previous page}} \\
%\toprule
%%\textbf{Patron} & \textbf{Archetype / Build Theme} & \textbf{Signature Mechanics} \\
%\midrule
%\endhead

%\midrule
%\multicolumn{3}{r}{{Continued on next page}} \\
%\endfoot
%\bottomrule
%\endlastfoot

%Carrion King & Decay Druid, Gentle Necromancer & Decay → renewal, rot as resource, harvest endings \\
%Clockwork Monad & Artificer, Engineer, Optimizer & Iterative improvement, temporary infusions, efficiency pressure \\
%Gallows Bell & Judge, Executioner, Notary & Oath enforcement, truth compulsion, final judgments \\
%Grimmir & Druid, Ranger, Hedge Witch & Seasonal sacrifice, primal magic, folk horror \\
%Ikasha & Shadow Monk, Spy, Dream Walker & Stealth, secrets, thresholds, raven omens \\
%Inaea & Hearthkeeper, Community Guardian, Healer & Webs of obligation, mercy as control, social binding \\
%Inquisitor Prime & Witch Hunter, Purifier, Zealot & Anti-magic, sanctified ground, banishment \\
%Isoka & Serpent Alchemist, Rebirth Cultist & Venom, shedding identity, transformation \\
%Khemesh & Warlock, Deep One, Abyssal Knight & Pressure, darkness, eldritch terror, crushing inevitability \\
%Livaea & Courtier, Seductress, Social Puppeteer & Charm, desire magic, emotional binding \\
%Lunara & Moon Priestess, Dreamwalker, Night Scholar & Revelation, thresholds, silver light, hidden ways \\
%Maelstraeus & Merchant Prince, Usurer, Contract Negotiator & Debt timers, infernal bargains, leverage economy \\
%Malachai & Cursed Champion, Faustian Bargainer & Cursed gifts, corruption, addictive power \\
%Morag & Hag, Trickster, Hedge Negotiator & Hidden costs, faerie bargains, knot magic \\
%Ninth & Archivist, Void Scholar, Pattern Seeker & Hyperknowledge, information overflow, paradox \\
%Oath of Flame \& Light & Paladin, Cleric, Dawn Knight & Vows, radiant smite, healing light \\
%Oath of Mercy and Grace & Hospitaller, Pacifist Healer, Forgiver & Unburdening, sanctuary, spiral mercy \\
%Pale Shepherd & Death Doula, Psychopomp, Grief Counsellor & Threshold guidance, melancholy wisdom, memory shepherding \\
%Palinode & Bard, Dancer, Performer & Unfinished art, audience binding, ecstatic performance \\
%Savage Heart & Berserker, Beastmaster, Primal Warrior & Predation, hunger, red maw \\
%Sealed Gate & Abjurer, Warden, Exorcist & Wards, thresholds, banishment, containment \\
%The Traveler & Guide, Scout, Pilgrim & Wayfinding, safe passage, road covenants \\
%Thrysos & Reveler, Satyr, Bacchanal & Ecstasy, frenzy, wine and dance \\
%Varnek Karn & Necromancer, Ancestor Speaker, Funeral Notary & Death bargains, ancestor consultation, unquiet host \\
%The Witness & Inquisitor, Archivist, Truth-Seeker & Revelation, memory, uncomfortable truths \\
%\end{longtable}
%
%\clearpage
%
%% --- Include the removed patron files (the ones that were commented out in the main chapter) ---
%\clearpage
%\input{patrons/gallows-bell}
%\clearpage
%\input{patrons/inquisitor-prime}
%\clearpage
%\input{patrons/khemesh}
%\clearpage
%\input{patrons/livaea}
%\clearpage
%\input{patrons/lunara}
%\clearpage
%\input{patrons/maelstraeus}
%\clearpage
%\input{patrons/malachai}
%\clearpage
%\input{patrons/morag}
%\clearpage
%\input{patrons/ninth-rim}
%\clearpage
%\input{patrons/palinode}
%\clearpage
%\input{patrons/sealed-gate}
%\clearpage
%\input{patrons/thrysos}
%\clearpage
%\input{patrons/varnek-karn}
%\clearpage

% --- Full 24-patron Gift Table (for completeness) ---
%\section{Complete Patron Gifts (All 24)}
%\label{app:full-gift-table}

%\centering
%\small
%\renewcommand{\arraystretch}{1.25}
%\setlength{\LTpre}{0.5em}
%\setlength{\LTpost}{0.5em}
%
%\begin{longtable}{@{}p{4.2cm}p{3.2cm}p{7.2cm}@{}}
%\toprule
%\textbf{Patron} & \textbf{Thematic Skill (+1)} & \textbf{Gift / Lore Bestowal} \\
%\midrule
%\endfirsthead
%
%\multicolumn{3}{c}{{\bfseries Continued from previous page}} \\
%\toprule
%\textbf{Patron} & \textbf{Thematic Skill (+1)} & \textbf{Gift / Lore Bestowal} \\
%\midrule
%\endhead
%
%\midrule
%\multicolumn{3}{r}{{Continued on next page}} \\
%\endfoot
%
%\bottomrule
%\endlastfoot
%
%Carrion King (Carrion, Renewal) & Survival &
%\emph{Grants the feast of decay, where what is dead becomes seed for what yet lives.} \\
%
%Clockwork Monad (Mechanism, Process) & Tinker &
%\emph{Grants the certainty of repetition: a cycle that never falters, a gear that never slips.} \\
%
%Gallows Bell (Doom, Last Rites) & Command &
%\emph{Grants the toll of ending, a voice that closes stories and calls all debts due.} \\
%
%Grimmir (Wild Speaker) & Survival &
%\emph{Grants the wisdom of root and stone, speaking the tongues of leaf and fang.} \\
%
%Ikasha (Shadow, Penumbra) & Stealth &
%\emph{Grants the hush between footsteps and the raven's omen at every threshold.} \\
%
%Inaea (Mercy, Hearth) & Persuade &
%\emph{Grants the hearth's warmth---shelter for the weary and mercy for the lost.} \\
%
%Inquisitor Prime (Purity, Persecution) & Insight &
%\emph{Grants the hunter's sight, piercing illusion and revealing supernatural corruption.} \\
%
%Isoka (Serpents, Shedding) & Skirmish &
%\emph{Grants the serpent's coil: strength in sudden strike and wisdom in renewal.} \\
%
%Khemesh (Abyss, Pressure) & Skirmish &
%\emph{Grants the crushing silence of the deep, where strength drowns beneath weight.} \\
%
%Livaea (Crimson Courtier) & Sway &
%\emph{Grants the velvet whisper, the smile that binds and the promise that cages.} \\
%
%Lunara (Silver Quiet) & Notice &
%\emph{Grants the silver mirror, the moonlit sight that pierces veils and hidden ways.} \\
%
%Maelstraeus (Commerce, Exchange) & Subterfuge &
%\emph{Grants the crooked scale, where every debt finds its true weight and every profit its hidden cost.} \\
%
%Malachai (Cruel Messenger) & Deception &
%\emph{Grants the honeyed curse, the gift that galls and the hope that withers.} \\
%
%Morag (Hag Bargains) & Sway &
%\emph{Grants the hidden cost, the crookèd thread where every gift has teeth and every kindness a snare.} \\
%
%Ninth (Hyperknowledge) & Lore &
%\emph{Grants the glimpse beyond, where patterns break and truth overflows the vessel of the mind.} \\
%
%Oath of Flame \& Light (Dawn, Vows) & Command &
%\emph{Grants the fire of dawn, a vow that shields the faithful and sears the faithless.} \\
%
%Oath of Mercy and Grace (Healing, Forgiveness) & Insight &
%\emph{Grants the spiral of mercy: every act of care pays forward, and the Ledger of Grace waits for no debtor. The Weepers' tears close pages that gold cannot touch.} \\
%
%Pale Shepherd (Thresholds, Guidance) & Stealth &
%\emph{Grants the shepherd's whisper, guiding the living through liminal spaces and comforting the dying.} \\
%
%Palinode (Performance, Rapture) & Performance &
%\emph{Grants the unfinished chorus, where art never dies and the encore is eternal.} \\
%
%Savage Heart (Primal Consumption) & Intimidation &
%\emph{Grants the red maw, the strength of bear and the hunger of the wolf.} \\
%
%Sealed Gate (Boundaries, Closure) & Tinker &
%\emph{Grants mastery of thresholds---doors that yield or bar at your command.} \\
%
%The Traveler (Ways, Roads) & Notice &
%\emph{Grants the open way, a restless compass and roads where none were marked.} \\
%
%Thrysos (Revels, Frenzy) & Performance &
%\emph{Grants the horned dance, the ecstasy that burns and the freedom that binds.} \\
%
%Varnek Karn (Ossuary, Dominion of the Dead) & Command &
%\emph{Grants the silence of the archive, where the dead obey and records speak.} \\
%
%The Witness (Truth, Revelation) & Notice &
%\emph{Grants the unblinking gaze that unmasks deceit and remembers every oath.} \\
%
%\end{longtable}
%%\par
%
%\captionof{table}{Complete Patron Gifts and Thematic Skills (24 patrons). The +1 bonus applies only when the fiction aligns with the Patron's domain.}

%\clearpage

%\section{Complete Patron Fear Influence Table (All 24)}
%\label{app:full-fear-table}
%
%\centering
%%\small
%\begin{longtable}{@{}p{3.2cm}p{4.5cm}p{2.8cm}p{5.9cm}@{}}
%%\toprule
%\textbf{Patron} & \textbf{Domain Tags} & \textbf{Influence Type} & \textbf{Fear Effect Modifications} \\
%\midrule
%\endfirsthead
%
%\multicolumn{4}{c}{{\bfseries Continued from previous page}} \\
%\toprule
%\textbf{Patron} & \textbf{Domain Tags} & \textbf{Influence Type} & \textbf{Fear Effect Modifications} \\
%\midrule
%\endhead
%
%\midrule
%\multicolumn{4}{r}{{Continued on next page}} \\
%\endfoot
%\bottomrule
%\endlastfoot
%
%Carrion King & Decay, Feast, Scavenging & Reroll & 
%Fear tagged \emph{Decay, Hunger, Feast} mutates. On 1 (Fight), frenzy devolves into cannibalistic hunger. \\
%
%Clockwork Monad & Order, Repetition, Mechanism & Amplify & 
%Fear tagged \emph{Order, Pattern, Repetition} becomes mechanistic horror. On 3 (Freeze), victim is trapped in endless loop. \\
%
%Gallows Bell & Doom, Last Rites & Resonance & 
%Fear tagged \emph{Ending, Toll, Justice} deepens. On 2 (Hide), victim hears the bell toll and accepts inevitable demise. \\
%
%Grimmir & Seasons, Verdict, Threshold & Reroll & 
%Fear tagged \emph{Wild, Nature, Season} becomes primal dread. On 6 (Submit), victim freezes as if rooted. \\
%
%Ikasha & Shadow, Whispers, Latent Potential & Amplify & 
%Fear tagged \emph{Shadow, Voice, Hidden} deepens. On 3 (Freeze), victim may be paralyzed by whispered secrets. \\
%
%Inaea & Mercy, Hearth, Connection & Reroll & 
%Fear tagged \emph{Home, Family, Comfort} twists to terror of loss. On 4 (Flee), victim abandons loved ones ``for their own good.'' \\

%Inquisitor Prime & Purity, Zeal, Persecution & Amplify & 
%Fear tagged \emph{Purity, Heresy, Persecution} turns fanatical. On 1 (Fight), victim attempts self-destructive martyrdom. \\
%
%Isoka & Serpents, Shedding, Venom & Reroll & 
%Fear tagged \emph{Change, Identity, Poison} becomes existential. On 5 (Hide), victim sheds their face and cannot be recognized. \\
%
%Khemesh & Abyss, Pressure, Depths & Amplify & 
%Fear tagged \emph{Depth, Pressure, Dark} crushes. On 1 (Fight), victim dives berserk into consuming dark. \\
%
%Livaea & Desire, Courts, Binding & Reroll & 
%Fear tagged \emph{Charm, Seduction, Betrayal} turns inward. On 2 (Hide), victim is overwhelmed by fear of intimacy or commitment. \\
%
%Lunara & Moon, Mirror, Veil & Resonance & 
%Fear tagged \emph{Secrets, Reflections, Hidden Ways} also distorts memory. On 5 (Hide), victim forgets why they were afraid. \\
%
%Maelstraeus & Debt, Commerce, Exchange & Amplify & 
%Fear tagged \emph{Wealth, Bargain, Trap} worsens. On 4 (Flee), victim drops coins or valuables in panic. \\
%
%Malachai & Curses, Corruption, False Hope & Reroll & 
%Fear tagged \emph{Curse, Poison, Broken Promise} deepens. On 1 (Fight), victim turns on a loved one, believing them to be the source of the curse. \\
%
%Morag & Bargains, Hidden Cost, Fate & Reroll & 
%Fear tagged \emph{Debt, Price, Promise} mutates. On 6 (Submit), victim accepts a deal they would otherwise refuse. \\
%
%Ninth & Knowledge, Paradox, Pattern & Amplify & 
%Fear tagged \emph{Truth, Logic, Revelation} becomes overwhelming. On 3 (Freeze), victim is lost in recursive thoughts. \\
%
%Oath of Flame \& Light & Dawn, Vows, Purity & Resonance & 
%Fear tagged \emph{Light, Sacred, Vow} also imposes Fatigue. Iconic light turns terror into bodily strain. \\
%
%Oath of Mercy and Grace & Healing, Forgiveness, Spiral & Reroll & 
%Fear tagged \emph{Care, Debt, Failure} becomes paralysing guilt. On 2 (Hide), victim flees from those they failed. \\
%
%Pale Shepherd & Thresholds, Souls, Passing & Reroll & 
%Fear tagged \emph{Death, Guidance, Soul} becomes existential. On 6 (Submit), victim follows willingly into afterlife path. \\
%
%Palinode & Performance, Rapture, Unfinished & Amplify & 
%Fear tagged \emph{Song, Audience, Failure} intensifies. On 4 (Flee), victim runs from imagined applause or rejection. \\
%
%Savage Heart & Predation, Hunger, Flesh & Reroll & 
%Fear tagged \emph{Beast, Blood, Prey} becomes primal. On 1 (Fight), victim charges the nearest target without thought. \\
%
%Sealed Gate & Boundaries, Closure & Amplify & 
%Fear tagged \emph{Confinement, Threshold, Trapped} worsens. On 4 (Flee), victim may injure themselves clawing at sealed barriers. \\
%
%The Traveler & Ways, Journeys, Crossroads & Resonance & 
%Fear tagged \emph{Lost, Path, Journey} also inflicts Complication Points. Terror of losing the way becomes mechanical hindrance. \\
%
%Thrysos & Revels, Frenzy, Excess & Reroll & 
%Fear tagged \emph{Dance, Wine, Madness} mutates. On 1 (Fight), victim enters a drunken frenzy. On 4 (Flee), they try to outrun their own shadow. \\
%
%Varnek Karn & Necromantic Archives, Death & Amplify & 
%Fear tagged \emph{Death, Undead, Necromancy} becomes horrific. On 2 (Hide), victim feigns death among corpses. \\
%
%The Witness & Truth, Revelation & Amplify & 
%Fear tagged \emph{Truth, Knowledge, Revelation} intensifies. On 6 (Submit), target may confess dark secrets under perceived Iconic judgment. \\
%
%\end{longtable}
