\chapter{Example Character Concepts}
łabel{ch:example−concepts}
∈dex{character concepts!examples}

This chapter presents example character concepts to illustrate how the game's systems can create diverse and interesting heroes. These are \textbf{examples only}−−−not prescriptive templates or exhaustive lists. Use them for inspiration, as pre−generated characters, or as starting points for your own unique creations.

\section{Important Disclaimer}
łabel{sec:concepts−disclaimer}
∈dex{character concepts!disclaimer}

\textbf{These examples are provided for illustrative purposes only.} They demonstrate how the game's mechanics can support different character archetypes and play styles. You are encouraged to:
\begin{itemize}
ℹtem Modify these concepts to fit your preferences
ℹtem Create completely original characters
ℹtem Mix and match elements from different examples
ℹtem Work with your Game Master to develop unique concepts
\end{itemize}

The game system is designed to support a wide variety of character types beyond these examples.

\section{How to Use These Examples}
łabel{sec:concepts−usage}
∈dex{character concepts!usage}

Each concept includes:
\begin{itemize}
ℹtem \textbf{Concept Overview}: Narrative identity and role
ℹtem \textbf{Traditional Analogues}: Common RPG classes that share similar themes −−− \textit{not} mechanical restrictions, just touchstones
ℹtem \textbf{Cultural Affinities}: Which regions or peoples commonly produce this archetype
ℹtem \textbf{Mechanical Foundation}: Suggested starting capabilities
ℹtem \textbf{Play Style}: How the character typically engages with challenges
ℹtem \textbf{Development Path (by Tier)}: What to buy at Tier I, II, and III (see \ref{sec:tiers})
ℹtem \textbf{Tactical Advice}: Tips for playing the role effectively
ℹtem \textbf{Story Hooks}: Plot opportunities for the Game Master
ℹtem \textbf{Legacy Note}: How this character might pass the torch (see \ref{ch:legacy−engine})
ℹtem \textbf{Build Blocks}: A \emph{32 XP} starting build, plus an optional \emph{36 XP} variant using Bonds/Complications (+4 XP) (see \ref{ch:advancement} for XP rules)
\end{itemize}

\section{1. The Guardian (Fighter / Paladin / Knight)}
łabel{sec:concept−guardian}
∈dex{character concepts!guardian}

\textbf{Concept}: A protector who stands between danger and those they've sworn to defend. \emph{Steel in hand, vow in heart.} Whether a chivalrous knight, a grizzled mercenary, or a holy warrior of the Everflame, the Guardian makes others safe through sacrifice.

\textbf{Traditional Analogues}: Fighter, Paladin, Barbarian (defensive), Knight, Samurai, Legionary

\textbf{Cultural Affinities}: Vhasia (chivalric knight), Viterra (hedge−knight), Aeler (shield−warden), Ubral (clan champion), Ecktoria (legionary), Oshiira (Vermilion Corps guardian)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Body (strength, endurance), Spirit (willpower, resolve)
ℹtem \textbf{Skills}: Melee, Athletics, Command, Endurance
ℹtem \textbf{Talents}: Defensive Survival, Combat Reflexes, Conditioning (see \ref{ch:talents})
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Frontline combat and protection
ℹtem Drawing attention away from allies (positioning, taunts, the Protect action)
ℹtem Using presence and authority to control situations
ℹtem Taking risks to protect others −−− absorbing Harm meant for teammates
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I (0−−40 XP)}: Raise Body to 3, Melee to 2, take \emph{Defensive Survival} and \emph{Combat Reflexes}. Use a shield.
ℹtem \textbf{Tier II (41−−90 XP)}: Raise Body to 4, take \emph{Conditioning} and \emph{Tactical Movement}. Consider a War Mount Asset (8 XP) for mobility.
ℹtem \textbf{Tier III (91−−150 XP)}: Take \emph{Battlefield Mastery} or \emph{Mystic Bulwark}. Upgrade armor to Heavy if you have Body 4+. Start training a successor (Cap 3+ follower).
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Use the Defend action} when outnumbered; it improves your Position by one step and can protect allies.
ℹtem \textbf{Spend Boons on re−rolling failed defense} rolls, not on offense. Your job is to survive.
ℹtem \textbf{Position matters more than damage.} A Dominant position lets you re−roll one failure −− use shoves, terrain, and cover to stay in control.
ℹtem \textbf{When you take Harm, remember it clears Fatigue.} Sometimes a small wound is worth the reset.
ℹtem \textbf{The Protect action} (intercepting a hit on an ally) is your signature move. Use it when a squishier ally is about to take a big hit.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem Who or what are they protecting? A person, a place, an ideal?
ℹtem What oath or duty drives them −−− sworn, inherited, or self−imposed?
ℹtem What happens if they fail in their protection?
ℹtem What personal costs do they bear for their role (scars, lost loved ones, broken oaths)?
ℹtem Did they inherit their duty from a fallen predecessor?
\end{itemize}

\textbf{Legacy Note}: A loyal follower (squire, apprentice, sworn shield) can become your successor, inheriting your sworn duty and reputation. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes} (Cost = rating \(×\) 3 XP): Body 3 (9), Spirit 2 (6), Wits 1 (3), Presence 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills} (Cost = level \(×\) 2 XP): Melee 2 (4), Athletics 1 (2), Command 1 (2) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Combat Reflexes (2 XP)
ℹtem Add \textbf{Talent}: Defensive Survival (3 XP)
ℹtem \textbf{Total}: 29 + 5 = 34 XP (bank 2 XP)
\end{itemize}

\section{2. The Mage (Wizard / Sorcerer / Arcanist)}
łabel{sec:concept−mage}
∈dex{character concepts!mage}

\textbf{Concept}: A wielder of arcane power who reshapes reality through study, intuition, or pact. \emph{Candlesmoke, marginalia, and dangerous truths.} From the rational Wizard of Thepyrgos to the instinctive Sorcerer of the Bloodlands, magic flows through them −−− and sometimes over them.

\textbf{Traditional Analogues}: Wizard, Sorcerer, Warlock (pact magic), Arcanist, Runekeeper, Invoker

\textbf{Cultural Affinities}: Thepyrgos (academic caster), Aelinnel (fae−touched), Valewood (Lethai−thora archivist), Ashaan (Breath−less resistance mage), Zakov (sea−witch), Sekogo (nganga spirit−worker)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Wits (theory, precision), Spirit (willpower, channeling)
ℹtem \textbf{Skills}: Lore, Arcana, Investigation, (optionally) Sway (for social casting)
ℹtem \textbf{Talents}: Spellcraft (freeform), Codex (runekeeper), Patron's Symbol (invoker)
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Information gathering and magical analysis
ℹtem Solving puzzles and mysteries with arcane insight
ℹtem Using magic to gain advantages in combat and social scenes
ℹtem Researching new spells and rituals between adventures
ℹtem Managing Influence (for Runekeepers and Invokers) or Backlash (for free casters)
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Take \emph{Spellcraft} (6 XP) or \emph{Codex} + \emph{Thiasos} (6 XP). Raise Wits to 3, Arcana/Lore to 2.
ℹtem \textbf{Tier II}: Add \emph{Elemental Focus} (4 XP) or \emph{Ritual Mastery} (4 XP). Raise Spirit to 3 for more Influence capacity.
ℹtem \textbf{Tier III}: Take \emph{Arcane Dominance} or \emph{Crack Specialist} (if Invoker). Build a library (Standard Asset, 8 XP) to research faster.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Push It} only when necessary −− extra Influence or Fatigue can snowball. Use Boons to improve Position before rolling.
ℹtem \textbf{Free casters:} Use cantrips (single tag) for safe, repeatable effects. Reserve multi−tag spells for critical moments.
ℹtem \textbf{Runekeepers:} Your Patron's Gift gives +1 Melee and +1 Thematic skill. Use it even if you aren't a fighter −− the Thematic skill can be Sway, Lore, or Stealth.
ℹtem \textbf{Invokers:} \emph{Crack the Seal} is for emergencies. Restore your Symbol during downtime before you need it again.
ℹtem \textbf{Learn the tags library.} Knowing which tags combine efficiently (e.g., AREA + BIND for crowd control) will make you a more effective caster.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What forbidden knowledge are they seeking?
ℹtem How did they acquire their magical talent −−− bloodline, study, pact, accident, trauma?
ℹtem What dangerous secrets might they uncover?
ℹtem Who opposes their research (rival mages, witch−hunters, the Church, the Chain−Lanterns)?
ℹtem What price have they already paid for power (a lost sense, a forgotten loved one, a cursed mark)?
\end{itemize}

\textbf{Legacy Note}: A trusted apprentice (follower) can inherit your Codex or Symbol collection, continuing your magical legacy. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Wits 3 (9), Spirit 2 (6), Body 1 (3), Presence 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Lore 2 (4), Arcana 1 (2) \(→\) \textbf{6 XP}
ℹtem \textbf{Total}: 27 XP (bank 3 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Spellcraft (6 XP) using banked 3 + 4 = 7 XP; bank 1 XP
ℹtem \textbf{Total}: 27 + 6 = 33 XP (bank 1 XP)
\end{itemize}

\section{3. The Rogue (Thief / Assassin / Scout)}
łabel{sec:concept−rogue}
∈dex{character concepts!rogue}

\textbf{Concept}: A shadow who moves where others cannot, taking what others cannot keep. \emph{Quiet footfalls, hawk eyes, and no loyalties but coin.} From the street−wise cutpurse of Silkstrand to the deadly assassin of Zakov, the Rogue survives by wits, speed, and one hidden knife.

\textbf{Traditional Analogues}: Rogue, Thief, Assassin, Scout, Ranger (stealth focus), Spy

\textbf{Cultural Affinities}: Zakov (pirate/rogue), Silkstrand (cutpurse), Vilikari (smuggler), Ykrul (steppe scout), Sidhi (lighthouse spy), Ashaan (Breath−less infiltrator)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Wits (perception, quick thinking), Body (agility, stealth)
ℹtem \textbf{Skills}: Stealth, Subterfuge, Investigation (or Streetwise), (optionally) Melee (light weapons)
ℹtem \textbf{Talents}: Backstab, Elusive Dodge, Shadow Dance (later −−− see \ref{ch:talents})
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Scouting ahead and gathering intelligence
ℹtem Infiltrating secured areas and disabling traps
ℹtem Ambush and skirmish tactics −−− striking from stealth
ℹtem Finding paths, contacts, and secrets in cities and wilderness
ℹtem Information brokering and black market connections
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Wits to 3, Stealth/Subterfuge to 2. Take \emph{Backstab} (6 XP) −−− you can afford it at 36 XP start.
ℹtem \textbf{Tier II}: Take \emph{Elusive Dodge} (6 XP). Raise Body to 3 for better tracking. Consider a Minor Asset (Hidden Cache, 4 XP).
ℹtem \textbf{Tier III}: Take \emph{Shadow Dance} (10 XP) and \emph{Deathblow} (12 XP) −−− the classic assassination chain. Build a spy network (Standard Asset, 8 XP).
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Always aim for Dominant Position before attacking from stealth.} Spend a Boon to improve Position if needed.
ℹtem \textbf{Use the environment.} Ask the GM: ``Is there a chandelier? A shadowed alcove? Loose gravel?'' Use it to gain advantage.
ℹtem \textbf{Backstab + Shadow Dance + Deathblow} is lethal, but requires Boons. Engage only when you have 2−−3 Boons saved.
ℹtem \textbf{When outmatched, retreat.} Spend a Boon to make a Controlled withdrawal instead of Desperate.
ℹtem \textbf{Your skills are not just for combat.} Use Subterfuge and Streetwise to navigate criminal networks, find black markets, and gather rumors.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What crime or secret forced them into the shadows?
ℹtem What treasure or truth are they hunting?
ℹtem How do they balance loyalty to allies with self−preservation?
ℹtem Who wants them dead −−− and why?
ℹtem What do they owe to the criminal underworld that trained them?
\end{itemize}

\textbf{Legacy Note}: A young pickpocket or informant (follower) could inherit your network of fences and contacts. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Wits 3 (9), Body 2 (6), Presence 1 (3), Spirit 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Stealth 2 (4), Subterfuge 2 (4) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Backstab (6 XP) −− needs 6 XP; bank 1 + 4 = 5 XP → not enough. Alternative: Add \textbf{Skill}: Investigation 1 (2 XP) and \textbf{Talent}: Combat Reflexes (2 XP) and bank 2 XP. Or take a Minor Asset: Hidden Cache (4 XP) and bank 2 XP.
ℹtem \textbf{Suggested}: Hidden Cache + bank 2 XP → total 33 XP.
\end{itemize}

\section{4. The Cleric (Priest / Healer / Exorcist)}
łabel{sec:concept−cleric}
∈dex{character concepts!cleric}

\textbf{Concept}: A vessel of divine or spiritual power, channeling healing, truth, and judgment. \emph{Sacred flame, holy water, and a prayer on the lips.} From the Lampers of Ecktoria to the fog−named exorcists of the Mistlands, the Cleric binds wounds and banishes darkness.

\textbf{Traditional Analogues}: Cleric, Priest, Paladin (healing focus), Exorcist, Shaman, Healer

\textbf{Cultural Affinities}: Ecktoria (Everflame priest), Viterra (Temple of Light), Thepyrgos (Covenant judge), Pereshi (Flame−Keeper), Tulkani (Caravan Judge), Sidhi (lighthouse healer)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Spirit (faith, willpower), Presence (authority, channeling)
ℹtem \textbf{Skills}: Lore (theology, rites), Medicine, Sway (prayer, exhortation), (optionally) Melee (symbolic weapon)
ℹtem \textbf{Talents}: Invoker (Patron's Symbol) or Runekeeper (Thiasos + Codex of a divine Patron like Everflame, Oath of Flame & Light, Pale Shepherd −− see \ref{ch:magic})
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Healing and protecting allies
ℹtem Purifying curses, diseases, and undead
ℹtem Ley line and ritual magic (wards, consecrations)
ℹtem Using authority of faith to command or persuade
ℹtem Exorcising spirits and banishing the Unquiet Dead
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Spirit to 3, Lore/Medicine to 2. Take a divine Patron's \emph{Symbol} (4 XP) or \emph{Codex} (4 XP).
ℹtem \textbf{Tier II}: Take \emph{Symbol Resonance} (4 XP) or \emph{Efficient Invocation} (4 XP). Raise Presence to 3 for better social influence.
ℹtem \textbf{Tier III}: Take \emph{Crack Specialist} or \emph{Patron's Favor}. Build a hospice or shrine (Standard Asset, 8 XP) to support your community.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Healing is powerful but slow.} In combat, stabilize Harm 2+ allies first; use HEAL (Fatigue recovery) on everyone else.
ℹtem \textbf{Your Patron's Symbol grants a free +1 to ritual rolls.} Keep it displayed.
ℹtem \textbf{Consecrate ground before a big fight} −− \emph{Rite of Consecrated Ground} can turn the tide against undead or demons.
ℹtem \textbf{Influence is story fuel.} When your Patron's presence is felt, answer −− it will lead to adventure.
ℹtem \textbf{Exorcism is negotiation, not combat.} Use Insight and Lore to understand what the spirit wants, then use Sway or Rites to resolve the haunting peacefully.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What crisis of faith have they survived?
ℹtem What heresy or supernatural threat haunts them?
ℹtem How do they serve both the living and the dead?
ℹtem What price do they pay for divine power?
ℹtem What secret does their faith hide from the faithful?
\end{itemize}

\textbf{Legacy Note}: An acolyte or lay healer (follower) can take up your Patron's Symbol and continue your mission. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Spirit 3 (9), Presence 2 (6), Wits 1 (3), Body 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Medicine 2 (4), Lore 1 (2), Sway 1 (2) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Patron's Symbol (4 XP) for a divine Patron (e.g., Oath of Flame & Light), using banked 1 + 4 = 5 XP; bank 1 XP
ℹtem \textbf{Total}: 33 XP (bank 1 XP)
\end{itemize}

\section{5. The Bard (Diplomat / Skald / Courtier)}
łabel{sec:concept−bard}
∈dex{character concepts!bard}

\textbf{Concept}: A master of words, music, and social influence who changes minds and opens doors. \emph{A smile for the foyer, steel for the parlor.} From the Skald of Linn to the Courtier of Vhasia, the Bard wields charm, rumor, and reputation as weapons.

\textbf{Traditional Analogues}: Bard, Diplomat, Courtier, Skald, Spy (social), Negotiator, Cantor

\textbf{Cultural Affinities}: Vhasia (courtier), Linn (skald), Thepyrgos (academic rhetorician), Tulkani (storyteller), Ecktoria (civic orator), Kahfagia (trade negotiator)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Presence (charm, performance), Wits (reading people, timing)
ℹtem \textbf{Skills}: Sway, Performance, Insight, (optionally) Lore (customs, history), Deception
ℹtem \textbf{Talents}: Silver Tongue, Inspire, Network Builder (or Read Emotions), Cantor's Path (optional, for Songs)
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Social interaction and negotiation
ℹtem Gathering information through rumors and contacts
ℹtem Resolving conflicts without violence (or making violence fashionable)
ℹtem Building alliances, manipulating factions
ℹtem Inspiring allies in combat (through Songs or simple words)
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Presence to 3, Sway to 2. Take \emph{Silver Tongue} (2 XP) and \emph{Inspire} (3 XP).
ℹtem \textbf{Tier II}: Take \emph{Command Presence} (4 XP) and \emph{Network Builder} (4 XP). Raise Wits to 3 for social reading.
ℹtem \textbf{Tier III}: Take \emph{High Stance Duelist} (if you use a blade) or \emph{Battlefield Terror}. Build a spy network (Standard Asset, 8 XP) or a courtly title (Major Asset, 12 XP). Consider taking \emph{Cantor's Path} (8 XP) for magical Songs.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Social scenes are your domain.} Use \emph{Inspire} to give allies +1 die or +1 Boon −− it can turn a negotiation.
ℹtem \textbf{Flashback declarations} (spend 1 Boon) are perfect for bards. ``I already bribed the guard last night.''
ℹtem \textbf{Don't neglect combat skills entirely.} A rapier and \emph{High Stance Duelist} can make you a credible threat.
ℹtem \textbf{Use \emph{Intricate} descriptions} to gain free re−rolls on 1s. Describe your performance, your charming smile, your family crest.
ℹtem \textbf{The best lie is built on truth.} Use Insight to learn what someone wants, then offer it to them (even if you have to fabricate it).
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What major conflict are they trying to resolve?
ℹtem What secret do they know that could topple a dynasty?
ℹtem How do they handle betrayal or failed negotiations?
ℹtem What personal relationships affect their diplomacy?
ℹtem Whose favor are they trying to win −−− and at what cost?
\end{itemize}

\textbf{Legacy Note}: A protégé or herald (follower) can inherit your reputation and contacts, keeping your network alive. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Presence 3 (9), Wits 2 (6), Spirit 1 (3), Body 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Sway 2 (4), Performance 1 (2), Insight 1 (2) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Inspire (3 XP) −−− once/scene, gain +1 die and potentially +1 Boon for bonded ally, using banked 1 + 4 = 5 XP; bank 2 XP
ℹtem \textbf{Total}: 32 XP (bank 2 XP)
\end{itemize}

\section{6. The Ranger (Hunter / Tracker / Beastmaster)}
łabel{sec:concept−ranger}
∈dex{character concepts!ranger}

\textbf{Concept}: A wilderness expert who navigates dangerous territories and hunts down quarry. \emph{Quiet footfalls, hawk eyes, and the long road.} From the Ykrul scout to the Valewood warden, the Ranger lives where roads end.

\textbf{Traditional Analogues}: Ranger, Hunter, Tracker, Beastmaster, Pathfinder, Wilderness Scout

\textbf{Cultural Affinities}: Ykrul (steppe scout), Kuvani (remount rider), Valewood (Lethai−al warden), Ubral (hill tracker), Mistlands (fog−walker), Sekogo (jungle hunter)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Wits (perception, navigation), Body (agility, endurance)
ℹtem \textbf{Skills}: Survival, Stealth, Nature (or Animal Handling), Archery or Melee
ℹtem \textbf{Talents}: Wilderness Lore, Keen Senses, (optionally) Animal Companion
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Scouting ahead and gathering intelligence
ℹtem Wilderness survival and navigation
ℹtem Tracking creatures or people across terrain
ℹtem Ambushing from concealment, using terrain
ℹtem (Optional) Bonding with an animal companion
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Wits to 3, Survival/Stealth to 2. Take \emph{Wilderness Lore} (2 XP) and \emph{Keen Senses} (2 XP).
ℹtem \textbf{Tier II}: Take \emph{Trackless Step} (2 XP) and \emph{Elusive Dodge} (6 XP) or a War Hound (Minor Asset, 4 XP). Raise Body to 3.
ℹtem \textbf{Tier III}: Take \emph{Shadow Dance} (10 XP) (if melee) or \emph{Relentless} (6 XP) for traversal. Upgrade your animal companion to a War Mount (Standard Asset, 8 XP) if allowed.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Use terrain to your advantage.} High ground, cover, and difficult terrain are your best friends. Ask the GM to set Position based on these.
ℹtem \textbf{Scout ahead} even when not required. Knowledge of what's coming negates ambushes and reduces travel timers.
ℹtem \textbf{If you have an animal companion}, use it to harass enemies (Distract & Draw independent action) rather than direct combat.
ℹtem \textbf{Survival 3+ means you rarely need to roll for basic tasks.} Leverage that for guaranteed successes in wilderness travel.
ℹtem \textbf{Know your prey.} Use Keen Senses to learn a creature's weakness before engaging.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What uncharted territory are they exploring?
ℹtem What secrets have they discovered in the wild?
ℹtem How do they balance civilization and wilderness?
ℹtem What prey (beast or man) still eludes them?
ℹtem What ancient guardian did they once encounter and survive?
\end{itemize}

\textbf{Legacy Note}: Your animal companion or a young tracker (follower) could carry on your legacy. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Wits 3 (9), Body 2 (6), Spirit 1 (3), Presence 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Survival 2 (4), Stealth 2 (4) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Keen Senses (2 XP)
ℹtem Add \textbf{Talent}: Wilderness Lore (2 XP)
ℹtem \textbf{Total}: 29 + 4 = 33 XP (bank 1 XP)
\end{itemize}

\section{7. The Survivor (Veteran / Outcast / Hardened)}
łabel{sec:concept−survivor}
∈dex{character concepts!survivor}

\textbf{Concept}: Someone who has endured hardship and developed resilience. \emph{Scars are maps; read them well.} From the war−torn mercenary to the plague survivor, the Survivor won't break −−− they've already been broken and rebuilt.

\textbf{Traditional Analogues}: Barbarian (resilience focus), Veteran, Ranger (endurance), Drifter, Refugee, Hollow−Touched

\textbf{Cultural Affinities}: Acasia (war survivor), Ubral (clan exile), Mistlands (fog−touched), Linn (shipwreck survivor), Zakov (pressed crew), Ashaan (escaped slave)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Spirit (willpower, resolve), Body (toughness)
ℹtem \textbf{Skills}: Endurance, Survival, (optionally) Perception (danger sense), Medicine (field)
ℹtem \textbf{Talents}: Iron Stomach, Conditioning, Second Wind (see \ref{ch:talents})
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Enduring difficult conditions (lack of food, exposure, poison)
ℹtem Overcoming physical and mental challenges through grit
ℹtem Using experience to avoid dangers
ℹtem Helping others survive hardships
ℹtem Taking hits that would fell others and still standing
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Spirit to 3, Endurance/Survival to 2. Take \emph{Iron Stomach} (3 XP) and \emph{Second Wind} (2 XP).
ℹtem \textbf{Tier II}: Take \emph{Conditioning} (4 XP) −− your Body counts as +1 for Fatigue track. Raise Body to 3.
ℹtem \textbf{Tier III}: Take \emph{Hollow−Touched} (8 XP) or \emph{Stone−Bound Interdictor} (8 XP) to convert Harm to Fatigue. Build a safehouse (Minor Asset, 4 XP) for respite.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Embrace the roller−coaster effect.} Taking Harm clears Fatigue −− sometimes a small wound is a tactical reset.
ℹtem \textbf{Iron Stomach makes you immune to mundane poisons and spoiled food.} Volunteer to test unknown rations.
ℹtem \textbf{Use \emph{Second Wind} to clear Fatigue before it overflows into Harm.}
ℹtem \textbf{When outmatched, use the environment.} Collapse a wall, start a fire, or trigger a rockfall. Your survival skills tell you how.
ℹtem \textbf{Known your limits, then push past them.} You can take more punishment than anyone else. Use that.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What trauma or hardship have they survived?
ℹtem How has their past shaped their present?
ℹtem What survival skills have saved them repeatedly?
ℹtem How do they help others facing similar challenges?
ℹtem What do they still fear, despite everything?
\end{itemize}

\textbf{Legacy Note}: A fellow survivor (follower) can inherit your network of caches and safehouses. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Spirit 3 (9), Body 2 (6), Wits 1 (3), Presence 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Endurance 2 (4), Survival 2 (4) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Iron Stomach (3 XP)
ℹtem \textbf{Total}: 29 + 3 = 32 XP (bank 2 XP)
\end{itemize}

\section{8. The Innovator (Artificer / Engineer / Tinker)}
łabel{sec:concept−innovator}
∈dex{character concepts!innovator}

\textbf{Concept}: A creative problem−solver who finds new solutions through craft and invention. \emph{Blueprints on napkins, tomorrow in your pocket.} From the Aeler engineer to the Gnome timermaker, the Innovator builds what others only dream.

\textbf{Traditional Analogues}: Artificer, Engineer, Tinker, Alchemist, Inventor, Savant

\textbf{Cultural Affinities}: Aeler (vent−prior), Aelinnel (clockwork artisan), Thepyrgos (academic engineer), Ngomebe (stone−singer), Cog−Cults (Monad devotee)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Wits (design, analysis), Body (fine manipulation)
ℹtem \textbf{Skills}: Craft, Tinker, Mechanics, (optionally) Alchemy, Investigation
ℹtem \textbf{Talents}: Quick Study, Technical Expert, (optionally) Magical Laboratory asset (see \ref{ch:assets−followers})
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Solving problems with unique inventions
ℹtem Creating or repairing specialized items
ℹtem Providing services others cannot (lockpicking, trap disarm, device creation)
ℹtem Using knowledge to gain advantages in planning
ℹtem Building and maintaining Assets (workshops, laboratories, caches)
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Wits to 3, Craft/Tinker to 2. Take \emph{Quick Study} (3 XP) and \emph{Workshop} (Minor Asset, 4 XP).
ℹtem \textbf{Tier II}: Take \emph{Efficient Invocation} (if magical) or \emph{Weapon Mastery} (if martial). Raise Body to 3 for fine manipulation.
ℹtem \textbf{Tier III}: Take \emph{Technical Expert} (custom) or \emph{Crack Specialist} (for magical items). Build a Guild Workshop (Standard Asset, 8 XP) to attract apprentices.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Preparation is your superpower.} Spend downtime building tools, traps, and contingencies. Your GM will reward that.
ℹtem \textbf{Use \emph{Quick Study} to learn enemy weaknesses} on the fly. ``I notice the golem's joints are unarmored −− aim there.''
ℹtem \textbf{Your workshop asset can jury−rig solutions.} Spend 1 Boon to ``have the right tool'' for a lock, a trap, or a repair.
ℹtem \textbf{When combat starts, use gadgets.} Caltrops, smoke bombs, and flash pellets can shift Position without a spell slot.
ℹtem \textbf{Document your inventions.} Keep a list of gadgets you've built and their effects −− this will make your play smoother and more creative.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What makes their specialty unique or valuable?
ℹtem How did they acquire their special skills (guild, apprenticeship, self−taught)?
ℹtem What problems require their specific expertise?
ℹtem Who seeks to control or exploit their abilities?
ℹtem What invention of theirs caused unintended harm?
\end{itemize}

\textbf{Legacy Note}: An apprentice (follower) can inherit your workshop and blueprints, continuing your inventive tradition. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Wits 3 (9), Body 2 (6), Presence 1 (3), Spirit 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Craft 2 (4), Tinker 2 (4) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Quick Study (3 XP) −−− +1 die to learn or adapt from observation, using banked 1 + 4 = 5 XP; bank 2 XP
ℹtem \textbf{Total}: 32 XP (bank 2 XP)
\end{itemize}

\section{9. The Summoner (Pact−Whisperer / Necromancer / Shaman)}
łabel{sec:concept−summoner}
∈dex{character concepts!summoner}

\textbf{Concept}: A binder of spirits, elementals, and otherworldly entities. \emph{Circles of salt, whispered names, and bargains struck in shadows.} From the necromancer of the Mistlands to the shaman of the Steppe, the Summoner commands powers that others fear.

\textbf{Traditional Analogues}: Necromancer, Conjurer, Summoner, Shaman, Spirit Caller, Demonologist

\textbf{Cultural Affinities}: Mistlands (bell−warden), Ykrul (bone−singer), Tulkani (dream−speaker), Ashaan (Breath−less spirit−binder), Zakov (drowned caller), Sekogo (nganga)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Spirit (willpower, binding), Wits (rituals, names)
ℹtem \textbf{Skills}: Arcana, Lore (spirits), (optionally) Command, Survival
ℹtem \textbf{Talents}: Pact−Whisperer (Lesser/Greater), Spirit Link, (optionally) Codex of Names (see \ref{ch:magic})
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Summoning creatures to fight, scout, or work
ℹtem Binding entities to solve problems (guard, research, transport)
ℹtem Negotiating with spirits and outsiders (social challenges)
ℹtem Managing Leash timers and Boon Finesse
ℹtem Building relationships with specific spirits over time (Spirit Bond progression)
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Take \emph{Pact−Whisperer} (2 XP) and \emph{Lesser Pactwright} (2 XP). Raise Spirit to 3, Arcana/Lore to 2.
ℹtem \textbf{Tier II}: Take \emph{Greater Pactwright} (4 XP) to summon Cap 3 spirits. Take \emph{Spirit Synergy} (4 XP) if you want two spirits.
ℹtem \textbf{Tier III}: Take \emph{Spirit Link} (10 XP) or \emph{True Name Keeper} (15 XP). Build a Spirit Circle (Minor Asset, 4 XP) to reduce Leash ticks.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Boon Finesse} is your most important tool. Spend 1 Boon to clear 1 Leash tick −− keep a reserve of 2−−3 Boons for emergencies.
ℹtem \textbf{Spirits are disposable but not free.} Use them for dangerous tasks (scouting a trapped room, triggering a trap) before risking your own skin.
ℹtem \textbf{When the Leash fills, the spirit acts once to its nature.} Plan for it −− position it near enemies before it breaks free.
ℹtem \textbf{Negotiate with spirits.} They have desires −− a promise of freedom, a drop of blood, a secret. A willing spirit ticks Leash slower.
ℹtem \textbf{Track Spirit Bonds.} If you summon the same spirit repeatedly, you can build a Spirit Bond Timer , reducing Leash ticks over time.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What entity are they bound to or seeking?
ℹtem What price did they pay for their first summoning?
ℹtem How do others react to their ``dark'' magic?
ℹtem Who hunts them for their forbidden practices?
ℹtem What debt do they owe to a spirit that saved their life?
\end{itemize}

\textbf{Legacy Note}: A bonded spirit or an apprentice medium (follower) could inherit your spirit contracts, continuing your work. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Spirit 3 (9), Wits 2 (6), Body 1 (3), Presence 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Arcana 2 (4), Lore 1 (2) \(→\) \textbf{6 XP}
ℹtem \textbf{Total}: 27 XP (bank 3 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Pact−Whisperer (2 XP)
ℹtem Add \textbf{Talent}: Lesser Pactwright (2 XP)
ℹtem \textbf{Total}: 27 + 4 = 31 XP (bank 3 XP; can add Familiar (2 XP) later or upgrade to Greater Pactwright)
\end{itemize}

\section{10. The Commander (Warlord / Tactician / Leader)}
łabel{sec:concept−commander}
∈dex{character concepts!commander}

\textbf{Concept}: A leader who directs allies, controls battlefields, and shapes conflicts through strategy. \emph{Orders in a whisper, victory in formation.} From the Condotta captain to the Ecktorian legate, the Commander wins before blades are drawn.

\textbf{Traditional Analogues}: Warlord, Marshal, Tactician, Commander, Strategist, Banneret

\textbf{Cultural Affinities}: Ecktoria (legate), Acasia (free company captain), Vhasia (marshal), Black Banners (condotta captain), Oshiira (Narrow Captain), Ykrul (war−leader)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Presence (authority, command), Wits (tactics, observation)
ℹtem \textbf{Skills}: Command, Tactics, (optionally) Sway (morale), Investigation (battlefield analysis)
ℹtem \textbf{Talents}: Inspire, Command Presence, Tactical Movement (for allies −− see \ref{ch:talents})
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Directing allies in combat (bonus dice, improved Position)
ℹtem Planning and executing tactical maneuvers
ℹtem Leading followers and assets in mass action (see \ref{ch:assets−followers})
ℹtem Using authority to intimidate or persuade
ℹtem Managing resources (supplies, morale, positioning) across extended operations
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Presence to 3, Command/Tactics to 2. Take \emph{Command Presence} (4 XP) and \emph{Inspire} (3 XP).
ℹtem \textbf{Tier II}: Take \emph{Tactical Movement} (4 XP) and \emph{Network Builder} (4 XP). Acquire a Cap 2−−3 follower (soldier, herald).
ℹtem \textbf{Tier III}: Take \emph{Battlefield Mastery} (6 XP) or \emph{Mystic Bulwark} (8 XP). Build a War Chest (Standard Asset, 8 XP) to fund operations.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Use \emph{Inspire} to give allies +1 die and +1 Boon.} A well−timed inspiration can turn a miss into a success.
ℹtem \textbf{Your followers are your strength.} Keep them alive −− spend Boons to protect them. A Cap 3 scout is worth more than a single attack.
ℹtem \textbf{Use \emph{Tactical Movement} to reposition without disengaging.} Order allies to flank while you hold the line.
ℹtem \textbf{In mass combat, treat your unit as a high−Cap follower.} Your Command roll can advance the War Timer faster than any sword.
ℹtem \textbf{Know when to fight and when to retreat.} A good commander wins battles; a great commander wins campaigns, sometimes by losing a skirmish.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What army or company did they lead before?
ℹtem What strategic defeat haunts them?
ℹtem Who are their rivals in the art of war?
ℹtem What cause do they fight for −−− coin, country, or creed?
ℹtem What soldiers died under their command, and how do they honor them?
\end{itemize}

\textbf{Legacy Note}: A loyal lieutenant or sergeant (follower) can inherit your command, leading the company after you. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Presence 3 (9), Wits 2 (6), Body 1 (3), Spirit 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Command 2 (4), Tactics 2 (4) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Command Presence (4 XP) −−− +1 die on leadership/intimidation rolls, using banked 1 + 4 = 5 XP; bank 1 XP
ℹtem \textbf{Total}: 33 XP (bank 1 XP)
\end{itemize}

\section{11. The Duelist (Swashbuckler / Fencer / Blade)}
łabel{sec:concept−duelist}
∈dex{character concepts!duelist}

\textbf{Concept}: A master of one−on−one combat, relying on speed, precision, and flair rather than heavy armor. \emph{Steel is a conversation; you speak it fluently.} From the Vhasian tournament knight to the Sidhi knife−dancer, the Duelist lives for the clash of blades.

\textbf{Traditional Analogues}: Swashbuckler, Fencer, Duelist, Blade, Rogue (combat focus), Monk (weapon focus)

\textbf{Cultural Affinities}: Vhasia (tourney knight), Viterra (legal duelist), Zakov (pirate duelist), Ubral (clan champion), Sidhi (knife−dancer), Nihori (wandering ronin)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Body (agility, speed), Wits (timing, reading opponents)
ℹtem \textbf{Skills}: Melee (Light weapons), Acrobatics, Perception, (optionally) Sway (taunt, feint)
ℹtem \textbf{Talents}: Weapon Master (Light Blade), Elusive Dodge, High Stance Duelist
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem One−on−one combat dominance
ℹtem Mobility and positioning (flanking, dancing around enemies)
ℹtem Using feints and taunts to control opponent's actions
ℹtem Parrying and riposting (converting defense into offense)
ℹtem Social dueling (fighting for honor or reputation, not just survival)
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Body to 3, Melee to 2. Take \emph{Weapon Master} (Light Blade, 5 XP) and \emph{Elusive Dodge} (2 XP).
ℹtem \textbf{Tier II}: Take \emph{High Stance Duelist} (6 XP) and \emph{Flurry Strike} (6 XP). Raise Wits to 3 for reading opponents.
ℹtem \textbf{Tier III}: Take \emph{Duelist's Edge} (6 XP) or \emph{Deathblow} (12 XP). Acquire a signature blade (Minor Asset, 4 XP) with an enchantment.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Your goal is to isolate a single target.} Use mobility to separate an enemy from the group.
ℹtem \textbf{Use \emph{High Stance Duelist} to gain advantage} when fighting a single opponent. Spend Boons to improve Position.
ℹtem \textbf{Flurry Strike lets you make two attacks} as one action. Use it to finish a wounded foe or to test a high−defense enemy.
ℹtem \textbf{Parry and riposte are your friends.} When you successfully defend, look for opportunities to counterattack as a free action (ask your GM).
ℹtem \textbf{Social dueling matters.} In honor cultures (Vhasia, Viterra, Ubral), winning a duel fairly can resolve conflicts without further bloodshed.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What duel did they lose that still stings?
ℹtem Whose honor are they sworn to defend?
ℹtem What signature technique did they develop themselves?
ℹtem Why did they leave their fencing school or tradition?
ℹtem What blade do they carry, and whose was it before theirs?
\end{itemize}

\textbf{Legacy Note}: A promising student (follower) can inherit your signature technique and blade. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Body 3 (9), Wits 2 (6), Spirit 1 (3), Presence 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Melee 2 (4), Acrobatics 1 (2), Perception 1 (2) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Weapon Master (Light Blade, 5 XP) using banked 1 + 4 = 5 XP; bank 0 XP
ℹtem \textbf{Total}: 34 XP (bank 0 XP)
\end{itemize}

\section{12. The Occultist (Investigator of the Unseen)}
łabel{sec:concept−occultist}
∈dex{character concepts!occultist}

\textbf{Concept}: A detective of the supernatural, solving mysteries that involve spirits, curses, and forbidden lore. \emph{Dusty tomes, silver mirrors, and the patience to listen to what does not speak.} From the Thepyrgos scholar to the Mistlands exorcist, the Occultist knows that every haunting has a witness.

\textbf{Traditional Analogues}: Investigator (supernatural), Occultist, Lorekeeper, Exorcist, Paranormal Detective, Witch−Finder

\textbf{Cultural Affinities}: Thepyrgos (academic occultist), Mistlands (bell−warden), Valewood (Lethai−thora archivist), Ashaan (Breath−less researcher), Tulkani (dream−interpreter), Sekogo (nganga diagnostician)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Wits (investigation, pattern recognition), Spirit (willpower, sensing the unseen)
ℹtem \textbf{Skills}: Investigation, Lore (spirits, curses), Insight, (optionally) Ritual
ℹtem \textbf{Talents}: Keen Senses, Occult Knowledge (custom), Ritual Mastery
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Investigating supernatural phenomena
ℹtem Identifying curses, possessions, and hauntings
ℹtem Performing exorcisms and banishing rites
ℹtem Researching forgotten lore in archives and libraries
ℹtem Consulting with spirits (through ritual or mediumship)
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Wits to 3, Investigation/Lore to 2. Take \emph{Keen Senses} (2 XP) and \emph{Occult Knowledge} (custom, 3 XP).
ℹtem \textbf{Tier II}: Take \emph{Ritual Mastery} (4 XP) and \emph{Contamination Control} (custom, 4 XP). Raise Spirit to 3.
ℹtem \textbf{Tier III}: Take \emph{Crack Specialist} or \emph{Hollow−Sight} (custom, 8 XP). Build an Archive (Standard Asset, 8 XP) to store your findings.
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Investigation is your primary tool.} Ask the GM questions: ``What kind of spirit is this? What unfinished business does it have? What's the anchor?''
ℹtem \textbf{Lore checks can reveal weaknesses.} A spirit may be vulnerable to a specific salt, a forgotten name, or an iron bell.
ℹtem \textbf{Rituals take time, but they work.} Use \emph{Ritual Mastery} to reduce casting time when the ghost is about to attack.
ℹtem \textbf{Document everything.} Your notes may become Assets (case files) that give bonuses to future investigations.
ℹtem \textbf{Sometimes the solution is not to banish, but to listen.} Many spirits just need to be witnessed.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What unsolved haunting drives them?
ℹtem What forbidden text did they read (and what did it cost them)?
ℹtem Who refuses to believe their findings?
ℹtem What spirit have they failed to help, and how does it haunt them?
ℹtem What secret does the Church want them to stop investigating?
\end{itemize}

\textbf{Legacy Note}: A student or junior investigator (follower) can inherit your case files and contacts. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Wits 3 (9), Spirit 2 (6), Body 1 (3), Presence 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Investigation 2 (4), Lore 2 (4) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Keen Senses (2 XP)
ℹtem Add \textbf{Talent}: Occult Knowledge (3 XP, custom: +1 die to identify supernatural creatures and curses)
ℹtem \textbf{Total}: 29 + 5 = 34 XP (bank 0 XP)
\end{itemize}

\section{13. The Infiltrator (Spy / Agent / Intelligence Officer)}
łabel{sec:concept−infiltrator}
∈dex{character concepts!infiltrator}

\textbf{Concept}: A master of disguise, deception, and covert action. \emph{Faces are costumes; trust is a vulnerability.} From the Vilikari spy to the Ashaani Breath−less agent, the Infiltrator moves through the world unnoticed, collecting secrets and planting lies.

\textbf{Traditional Analogues}: Spy, Agent, Infiltrator, Social Rogue, Intelligence Officer, Saboteur

\textbf{Cultural Affinities}: Vilikari (border agent), Ashaan (Breath−less spy), Ecktoria (Censorate informer), Thepyrgos (Chain−Lantern mole), Silkstrand (information broker), Zakov (syndicate agent)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Presence (disguise, deception), Wits (observation, planning)
ℹtem \textbf{Skills}: Deception, Subterfuge, Insight, (optionally) Stealth
ℹtem \textbf{Talents}: Silver Tongue, Silk Palms, Shadow Dance (for escapes)
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Infiltrating secure locations under false identities
ℹtem Planting misinformation and framing rivals
ℹtem Extracting secrets from unwilling targets
ℹtem Operating behind enemy lines
ℹtem Maintaining a cover identity for months or years
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Presence to 3, Deception/Subterfuge to 2. Take \emph{Silver Tongue} (2 XP) and \emph{Silk Palms} (2 XP).
ℹtem \textbf{Tier II}: Take \emph{Trackless Step} (2 XP) and \emph{Network Builder} (4 XP). Raise Wits to 3.
ℹtem \textbf{Tier III}: Take \emph{Shadow Dance} (10 XP) for escapes. Build a Spy Network (Standard Asset, 8 XP).
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Cover identities are your most valuable asset.} Spend downtime building them. Have at least one backup.
ℹtem \textbf{Use \emph{Silk Palms} to create temporary Strings.} A bribe today becomes a favor tomorrow.
ℹtem \textbf{Never reveal your true agenda.} Let your targets think you're a merchant, a pilgrim, a lost traveler.
ℹtem \textbf{Always have an escape route.} \emph{Shadow Dance} is expensive, but worth it when the guard patrol turns the corner.
ℹtem \textbf{The best intelligence is what they give you willingly.} Use Insight to learn what someone wants, then position yourself as the one who can provide it.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What intelligence agency or patron do they serve?
ℹtem What secret are they trying to uncover?
ℹtem What cover identity has become uncomfortably real?
ℹtem Who burned them, and how will they get even?
ℹtem What would they sacrifice to protect their true identity?
\end{itemize}

\textbf{Legacy Note}: A handler or junior agent (follower) can inherit your network of covers and contacts. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Presence 3 (9), Wits 2 (6), Body 1 (3), Spirit 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Deception 2 (4), Subterfuge 2 (4) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Silver Tongue (2 XP)
ℹtem Add \textbf{Talent}: Silk Palms (2 XP)
ℹtem \textbf{Total}: 29 + 4 = 33 XP (bank 1 XP)
\end{itemize}

\section{14. The Channeler (Medium / Spiritualist / Oracle)}
łabel{sec:concept−channeler}
∈dex{character concepts!channeler}

\textbf{Concept}: A vessel for spirits, voices, and visions from beyond the veil. \emph{The dead speak through me; I am their mouth and their mercy.} From the Tulkani dream−speaker to the Mistlands spirit−medium, the Channeler bridges the living and the dead.

\textbf{Traditional Analogues}: Medium, Spiritualist, Oracle, Seer, Shaman (spirit channeling), Vessel

\textbf{Cultural Affinities}: Tulkani (dream−speaker), Mistlands (spirit−medium), Sekogo (nganga channeler), Theona (Green Host whisperer), Ashaan (Breath−less spirit−binder), Linn (volva)

\textbf{Mechanical Foundation}:
\begin{itemize}
ℹtem \textbf{Primary}: Spirit (willpower, channeling), Presence (conduit strength)
ℹtem \textbf{Skills}: Ritual, Lore (spirits), Insight, (optionally) Performance (trance work)
ℹtem \textbf{Talents}: Summoning (spirit focus), Cantor's Path (for spirit−Songs), Spirit Link
\end{itemize}

\textbf{Play Style}:
\begin{itemize}
ℹtem Channeling spirits to gain knowledge or perform tasks
ℹtem Serving as a medium between the living and the dead
ℹtem Receiving visions and prophecies (with GM guidance)
ℹtem Performing seances and spirit consultations
ℹtem Managing the risk of possession or spiritual contamination
\end{itemize}

\paragraph{Development Path (by Tier)}:
\begin{itemize}
ℹtem \textbf{Tier I}: Raise Spirit to 3, Ritual/Lore to 2. Take \emph{Pact−Whisperer} (2 XP) and a spiritual Patron (Ikasha, Pale Shepherd, or Witness).
ℹtem \textbf{Tier II}: Take \emph{Spirit Link} (10 XP) or \emph{Cantor's Path} (8 XP). Raise Presence to 3.
ℹtem \textbf{Tier III}: Take \emph{True Name Keeper} (15 XP) or \emph{Embraced Corruption} (12 XP for Cantors). Build a Spirit Circle (Standard Asset, 8 XP).
\end{itemize}

\paragraph{Tactical Advice}:
\begin{itemize}
ℹtem \textbf{Channeling is risky.} Each time you let a spirit speak through you, tick a Spirit Influence timer. On fill, the spirit may try to stay.
ℹtem \textbf{Visions are clues, not scripts.} When you receive a prophecy, work with your GM to interpret it. It may be cryptic, but it's never false.
ℹtem \textbf{Use Ritual to consult the dead.} A successful ritual can answer one question about a death or a hidden secret.
ℹtem \textbf{Protect yourself.} Salt circles, iron bells, and witness cords can keep unwanted spirits from piggybacking on your channeling.
ℹtem \textbf{The dead have their own agendas.} A spirit may answer your question, but it may also ask for something in return.
\end{itemize}

\textbf{Story Hooks}:
\begin{itemize}
ℹtem What spirit first spoke through them, and why?
ℹtem What prophecy do they carry that no one believes?
ℹtem How do they manage the constant presence of the dead?
ℹtem What spirit wants to possess them permanently, and how do they resist?
ℹtem What secret do the dead want revealed, and who wants it to stay hidden?
\end{itemize}

\textbf{Legacy Note}: A student medium (follower) can inherit your spirit contacts and channeling techniques. See \ref{ch:legacy−engine}.

\paragraph{Build Blocks.}
\textbf{Starting Build (32 XP).}
\begin{itemize}
ℹtem \textbf{Attributes}: Spirit 3 (9), Presence 2 (6), Wits 1 (3), Body 1 (3) \(→\) \textbf{21 XP}
ℹtem \textbf{Skills}: Ritual 2 (4), Lore 1 (2), Insight 1 (2) \(→\) \textbf{8 XP}
ℹtem \textbf{Total}: 29 XP (bank 1 XP)
\end{itemize}
\textbf{With Bonds/Complications (36 XP).}
\begin{itemize}
ℹtem Add \textbf{Talent}: Pact−Whisperer (2 XP)
ℹtem \textbf{Total}: 31 XP (bank 3 XP; can add Cantor's Path or Spirit Link later)
\end{itemize}

\section{Reskinning the Paths: Runekeeper and Invoker as Familiar Concepts}
łabel{sec:reskinning−paths}
∈dex{reskinning}∈dex{Runekeeper!reskinning}∈dex{Invoker!reskinning}

The magic systems in \textit{Fate's Edge} are designed to be reimagined. Neither \textbf{Runekeeper} (Thiasos + Codex) nor \textbf{Invoker} (Patron's Symbol) is locked into a single narrative identity. Below are common reskins −−− use the same mechanics, change the fiction.

\subsection*{Runekeeper −− The Devoted Agent}
A Runekeeper has sworn themselves to a single Patron, learning its Rites and bearing its Influence. This framework can represent:

\begin{itemize}
  ℹtem \textbf{Paladin / Crusader}: Sworn to the \textit{Oath of Flame & Light} or the \textit{Inquisitor Prime}. Rites become prayers, smites, and consecrations. The Codex is a prayer book or a vow−etched gauntlet.
  ℹtem \textbf{Warlock / Pact−Bearer}: Bound to a distant, alien, or dangerous Patron (e.g., Khemesh, Malachai, the Ninth). Rites are granted, not learned. Influence is the steady pull of a bargain.
  ℹtem \textbf{Cleric / Priest}: Devoted to the Everflame, the Pale Shepherd, or other divine powers. Rites are liturgies. The Patron's Gift is a blessing or a relic.
  ℹtem \textbf{Dark Knight / Death Knight}: A fallen or morally ambiguous warrior who draws power from a grim Patron (Carrion−King, Mor'iraath). Melee and Rites mix; Influence is a slow doom.
\end{itemize}

\noindent \textit{Mechanical core unchanged:} 1 action Rites, Influence track, Patron's Gift imbuement, Push It option.

\subsection*{Invoker −− The Professional Occultist}
An Invoker does not fully commit to a single Patron. Instead, they carry \textbf{Symbols} (often tools, badges, or relics) of several Powers and perform rituals to invoke their Rites. This framework can represent:

\begin{itemize}
  ℹtem \textbf{Exorcist / Witch Hunter}: Carries symbols of several protective Patrons (The Sealed Gate, Oath of Flame, Gallow's Bell). Rites are wards, banishments, and purification rituals. \emph{Crack the Seal} is an emergency exorcism.
  ℹtem \textbf{Alienist / Scholar of the Unseen}: Studies forbidden Patrons (The Ninth, Khemesh, Mor'iraath) from a safe distance −−− ``safe'' being relative. Uses Symbol libraries to invoke their power without full devotion. Rituals are experimental and dangerous.
  ℹtem \textbf{Consulting Occultist / Paranormal Investigator}: Carries a briefcase or satchel of Symbols −−− a bell from the Mistlands, a sun−seal from Ecktoria, a hag's stone from Aelinnel. Hired to solve magical problems. Rituals are filed like work orders; \emph{Crack the Seal} is ``burning a favor.''
  ℹtem \textbf{Ritualist / Ceremonial Mage}: Believes that power lies in procedure, not faith. Each Symbol is a ``key'' −−− a geometric diagram, a calendar, a consecrated blade. Rites are slow, deliberate, and precise. The Invoker is a technician, not a worshipper.
\end{itemize}

\noindent \textit{Mechanical core unchanged:} Ritual time, Influence per invocation, Symbol display reduces cost, \emph{Crack the Seal} for instant effect.

\subsection*{Mixing and Blurring Lines}
Characters may also blend these frameworks −−− a Runekeeper who occasionally carries a second Symbol for emergency rites, or an Invoker who one day pledges fully to a Patron and becomes a Runekeeper. The system does not penalize narrative growth; simply retrain Talents during downtime (see \ref{ch:talents}) when the fiction justifies it.

\noindent \textit{Remember:} The fiction leads. If you want to play a Paladin, describe your Runekeeper's Rites as prayers. If you want a Vodouisant or Hoodoo practitioner, reflavor Patron's Symbols as sacred bundles, veves, or spirit vessels. If you want a monster hunter, carry a satchel of monster−specific Symbols (silver stake, iron bell, cold iron nail). The mechanics stay the same; the story breathes.

\section{Creating Your Own Concept}
łabel{sec:creating−own−concept}
∈dex{character concepts!creation}

\subsection*{Start with Narrative}
\begin{itemize}
ℹtem What is your character's background and motivation?
ℹtem What role do they play in their community or society?
ℹtem What relationships are important to them?
ℹtem What goals are they pursuing?
\end{itemize}

\subsection*{Add Mechanical Support}
\begin{itemize}
ℹtem Choose attributes that support your concept (see \ref{sec:core−attributes})
ℹtem Select skills that reflect their training and experience (see \ref{sec:skills})
ℹtem Consider talents that provide unique capabilities (see \ref{ch:talents})
ℹtem Think about assets that represent their resources (see \ref{ch:assets−followers})
\end{itemize}

\subsection*{Consider Group Role}
\begin{itemize}
ℹtem How does your concept complement other party members?
ℹtem What gaps in group capability can you fill?
ℹtem What unique contributions can you make?
ℹtem How will you work with other characters?
\end{itemize}

\subsection*{Plan for Growth}
\begin{itemize}
ℹtem What short−term improvements make sense?
ℹtem What long−term development aligns with your concept?
ℹtem How might your character change over time?
ℹtem What legacy do you want to build? (See \ref{ch:legacy−engine})
\end{itemize}

\begin{tcolorbox}[colback=green!5!white,colframe=green!75!black,title=Character Concept Worksheet,fonttitle=\bfseries]
∈dex{character worksheet}
\textbf{Narrative Elements:}
\begin{itemize}
ℹtem Concept: ______________________
ℹtem Traditional Analogues: ________________
ℹtem Cultural Origin: ______________________
ℹtem Motivation: ______________________
ℹtem Background: ______________________
ℹtem Relationships: ______________________
\end{itemize}

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary Attributes: ______________________
ℹtem Key Skills: ______________________
ℹtem Starting Talents: ______________________
ℹtem Initial Assets: ______________________
\end{itemize}

\textbf{Development Plan:}
\begin{itemize}
ℹtem Short−term goals: ______________________
ℹtem Long−term vision: ______________________
\end{itemize}
\end{tcolorbox}

\section{Final Notes}
łabel{sec:concepts−final}
∈dex{character concepts!final notes}

Remember that these examples are starting points, not limitations. The most interesting characters often combine elements from multiple concepts or create entirely new approaches. Work with your Game Master to ensure your character concept fits the campaign and provides engaging story opportunities.

The best characters are those that you find interesting to play and that contribute to an enjoyable experience for everyone at the table.

\subsection*{Quick Reference: Concept Summary Table}

\begin{center}
\begin{longtable}{p{2.5cm}p{2.5cm}p{8cm}}
⊤rule
\textbf{Concept} & \textbf{Primary Attributes} & \textbf{Key Talents} \\
∣rule
Guardian & Body, Spirit & Defensive Survival, Combat Reflexes, Conditioning \\
Mage & Wits, Spirit & Spellcraft, Codex, Patron's Symbol \\
Rogue & Wits, Body & Backstab, Elusive Dodge, Shadow Dance \\
Cleric & Spirit, Presence & Patron's Symbol, Ritual Mastery, Symbol Resonance \\
Bard & Presence, Wits & Silver Tongue, Inspire, Network Builder \\
Ranger & Wits, Body & Wilderness Lore, Keen Senses, Trackless Step \\
Survivor & Spirit, Body & Iron Stomach, Conditioning, Second Wind \\
Innovator & Wits, Body & Quick Study, Workshop, Technical Expert \\
Summoner & Spirit, Wits & Pact−Whisperer, Greater Pactwright, Spirit Link \\
Commander & Presence, Wits & Command Presence, Inspire, Tactical Movement \\
Duelist & Body, Wits & Weapon Master, High Stance Duelist, Flurry Strike \\
Occultist & Wits, Spirit & Keen Senses, Occult Knowledge, Ritual Mastery \\
Infiltrator & Presence, Wits & Silver Tongue, Silk Palms, Shadow Dance \\
Channeler & Spirit, Presence & Pact−Whisperer, Spirit Link, Cantor's Path \\
⊥tomrule
\end{longtable}
\end{center}