\chapter{Assets and Followers}
\label{ch:assets-followers}
\index{assets}\index{followers}

Power in Fate's Edge is not measured only by what your character can do with their own hands.  
As characters grow, their influence extends outward---into places, people, and institutions that act in their name.

\textbf{Assets} and \textbf{Followers} represent this outward reach.  
They are how a character's past choices continue to matter when the dice stop rolling.

\emph{Assets reshape the world around you.}  
\emph{Followers stand beside you when the world pushes back.}

\begin{tcolorbox}[colback=green!5!white,colframe=green!75!black,title=A Word from the Raven Road]
\emph{"A safehouse is not a home; it is a promise you may not keep. A follower is not a servant; they are a weight you choose to carry. Invest wisely, or the road will collect."}  
\vspace{0.2cm}\hfill --- Dusana of the Raven Road, \textit{The Hearth Ledger}
\end{tcolorbox}

\section{Understanding Assets and Followers}
\label{sec:assets-followers-overview}
\index{assets!overview}\index{followers!overview}

Assets and Followers serve different roles in play, but both are expressions of legacy, leverage, and risk.

\subsection*{Key Differences}
\begin{itemize}
\item \textbf{Assets} are off‑screen resources: places, networks, titles, or enterprises.
\item \textbf{Followers} are on‑screen allies who act during scenes.
\item \textbf{Assets} change the situation \emph{before or around} a scene.
\item \textbf{Followers} participate \emph{directly} in the moment.
\end{itemize}

Both are powerful---but neither is free.

\subsection*{Management and Risk}
Assets and Followers require care and invite consequences:
\begin{itemize}
\item They demand upkeep through XP, downtime, or narrative obligation.
\item They can be targeted, compromised, or destroyed.
\item Overuse draws attention from enemies, factions, and rivals.
\end{itemize}

\emph{Power creates gravity. What you rely on can be taken from you.}

\section{Assets System}
\label{sec:assets-system}
\index{assets!system}

Assets are persistent resources under your control: holdings, institutions, reputations, or infrastructures.

\textit{\textbf{Assets do not replace rolls. They change what rolls are possible.}}

Assets allow characters to:
\begin{itemize}
\item Arrive prepared instead of improvising.
\item Reframe danger into opportunity.
\item Shift Position without negating risk.
\end{itemize}

\subsection*{Asset Types and Costs}
\label{sec:asset-types-costs}
\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Type} & \textbf{XP Cost} & \textbf{Establishment Time} \\
\midrule
Minor    & 4 XP  & 1 day \\
Standard & 8 XP  & 1 week \\
Major    & 12 XP & 1 month \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Asset Examples}
\label{sec:asset-examples}
\begin{description}
\item[Minor Assets] Safehouse, small shop, minor title, basic workshop.
\item[Standard Assets] Guild membership, noble office, trading post, spy network.
\item[Major Assets] Fortress, chartered city license, major enterprise, regional authority.
\end{description}

\subsection*{Using Assets}
\label{sec:using-assets}

Assets provide leverage in two primary ways: quiet preparation and dramatic activation.

\paragraph{Free Off‑Screen Use}
\label{sec:asset-free-use}
Each Asset has a defined off‑screen effect usable \textbf{once per session}, typically during downtime or scene framing:
\begin{itemize}
\item \textbf{Safehouse}: Secure lodging or concealment.
\item \textbf{Spy Network}: Baseline intelligence about a person or place.
\item \textbf{Workshop}: Repair gear or prepare simple solutions.
\item \textbf{Trading Post}: Acquire common goods or establish contacts.
\end{itemize}
These uses prepare the ground---they do not resolve active conflict.

\paragraph{Boon Activation}
\label{sec:asset-boon-activation}
Spend \textbf{1 Boon} (see \ref{ch:core-mechanics}) to activate an Asset dramatically during a scene.

This does not guarantee success. Instead, the Asset \textbf{reshapes the fiction} so a different outcome becomes possible.

An activated Asset may:
\begin{itemize}
  \item Improve \textbf{Position} by one step.
  \item Reduce the effective \textbf{DV} of a follow‑up action.
  \item Advance, suppress, or redirect a relevant Clock.
  \item Grant temporary access, protection, or narrative permission.
\end{itemize}

\textbf{Examples:}
\begin{itemize}
  \item \textbf{Safehouse}: Reveal a concealed exit, allowing escape at \emph{Controlled} Position.
  \item \textbf{Spy Network}: Surface critical intelligence, advancing an investigation Clock.
  \item \textbf{Workshop}: Jury‑rig a solution that enables an otherwise impossible attempt.
  \item \textbf{Trading Post}: Call in a favor to gain leverage or access in a negotiation.
\end{itemize}

The GM determines the precise effect based on the Asset's scale, condition, and the current stakes.

\paragraph{High‑Tier Use}
\label{sec:asset-high-tier}
At higher tiers (see \ref{sec:tiers}), Assets carry greater narrative weight---and greater risk.

Repeated or aggressive use may:
\begin{itemize}
  \item Create new complications.
  \item Advance hostile faction Clocks.
  \item Place the Asset itself in danger.
\end{itemize}
\emph{Assets are leverage, not immunity.}

\paragraph{XP Activation}
\label{sec:asset-xp-activation}
Spend \textbf{2 XP} (see \ref{ch:advancement}) to use an Asset's off‑screen effect beyond its normal limit:
\begin{itemize}
\item Emergency use after the free activation is spent.
\item Additional downtime applications.
\item Exceptional circumstances requiring extended support.
\end{itemize}
XP use represents strain, overreach, or long‑term cost.

\section{Asset Conditions}
\label{sec:asset-conditions}
\index{assets!conditions}

Assets have condition states affecting their usefulness.

\subsection*{Condition Levels}
\label{sec:asset-condition-levels}
\begin{description}
\item[Maintained] Fully functional, no penalties.
\item[Neglected] --1 die when used; requires attention.
\item[Compromised] Unavailable until repaired or recovered.
\end{description}

\subsection*{Maintenance Requirements}
\label{sec:asset-maintenance}
\begin{itemize}
\item \textbf{Regular Upkeep}: Two options (see \ref{subsec:upkeep-options}).
\item \textbf{Neglect}: Assets deteriorate if not maintained.
\item \textbf{Recovery}: Compromised assets require significant effort to restore.
\end{itemize}

\section{Followers System}
\label{sec:followers-system}
\index{followers!system}

Followers are characters who assist you directly.

\subsection*{Follower Capability Ratings}
\label{sec:follower-cap}
Followers are rated by Capability (\textbf{Cap}) from 1 to 5:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Cap} & \textbf{Description} \\
\midrule
1 & Novice helper, basic assistance \\
2 & Competent assistant, reliable support \\
3 & Skilled specialist, valuable aid \\
4 & Expert ally, significant capability \\
5 & Master companion, exceptional ability \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Follower Costs}
\label{sec:follower-costs}
\begin{itemize}
\item \textbf{XP Cost}: Capability squared ($\text{Cap}^2$).
\item \textbf{Example}: Cap 3 follower costs $3^2 = 9$ XP.
\item \textbf{Recruitment}: 1--3 days downtime to find and brief.
\item \textbf{Limits}: The GM may set maximum followers based on story.
\end{itemize}

\subsection*{Follower Types}
\label{sec:follower-types}
\begin{description}
\item[Combat Allies] Warriors, guards, mercenaries.
\item[Technical Experts] Craftspeople, engineers, specialists.
\item[Social Contacts] Informants, diplomats, agents.
\item[Specialists] Unique capabilities like magic or stealth.
\end{description}

\section{Using Followers}
\label{sec:using-followers}
\index{followers!usage}

\subsection*{Assistance in Scenes}
\label{sec:follower-assistance}
Followers can help with your actions:
\begin{itemize}
\item \textbf{Assist Dice}: Add dice equal to $\min(\text{Cap}, \text{relevant skill})$.
\item \textbf{Maximum Bonus}: +3 dice total from all sources (see \ref{ch:core-mechanics}).
\item \textbf{Cost}: Spend 1 Boon or 1 Fatigue to add +1 die (max +3 from assists).
\item \textbf{One Helper}: Only one follower can assist per action.
\end{itemize}

\subsection*{Independent Actions}
\label{sec:follower-independent}
Once per scene (party‑wide), a follower can take a small action:
\begin{description}
\item[\emph{Scout \& Signal}] Change an ally's next action to \textbf{Dominant} position.
\item[\emph{Distract \& Draw}] Reduce a threat clock by 1 segment.
\item[\emph{Fetch \& Carry}] Move an object through danger safely.
\end{description}

\subsection*{Cost of Independent Actions}
\label{sec:follower-independent-cost}
\begin{itemize}
\item Mark +1 \textbf{Exposure} (attention or stress), \emph{or}
\item Take \textbf{Harm 1} (injury or trauma).
\item Cannot be used if the follower is already \textbf{Compromised}.
\end{itemize}

\section{Follower Conditions}
\label{sec:follower-conditions}
\index{followers!conditions}

Followers track two condition types.

\subsection*{Exposure}
\label{sec:follower-exposure}
Represents attention, stress, or narrative pressure:
\begin{itemize}
\item \textbf{Gains}: From independent actions, dangerous situations, complications.
\item \textbf{Effects}: Increased risk, reduced effectiveness, attention from enemies.
\item \textbf{Recovery}: Downtime activities, careful management.
\end{itemize}

\subsection*{Harm}
\label{sec:follower-harm}
Represents injury, trauma, or damage:
\begin{itemize}
\item \textbf{Gains}: From combat, accidents, enemy attacks.
\item \textbf{Effects}: Penalties to assistance, possible incapacity.
\item \textbf{Recovery}: Medical care, rest, magical healing.
\end{itemize}

\subsection*{Condition States}
\label{sec:follower-condition-states}
\begin{description}
\item[Maintained] Ready and reliable, full capability.
\item[Neglected] Needs attention, --1 die to assistance.
\item[Compromised] Unavailable: captured, defected, lost, or incapacitated.
\end{description}

\section{Follower Risks}
\label{sec:follower-risks}
\index{followers!risks}

Using followers carries significant risks.

\subsection*{Complication Targeting}
\label{sec:follower-complication}
When the GM spends \textbf{2+ Story Beats} (see \ref{ch:core-mechanics}) on an action where you have assistance:
\begin{itemize}
\item The follower may face consequences instead of you.
\item Consequences could be injury, capture, betrayal, or other complications.
\item The outcome must be fictionally appropriate.
\end{itemize}

\subsection*{Off‑Screen Capability}
\label{sec:follower-offscreen}
Once per downtime, a \textbf{Cap 5} follower can solve a significant problem:
\begin{itemize}
\item But generates \textbf{1 Story Beat} for the party.
\item The GM describes how this creates story consequences.
\item Useful for emergencies, but costly.
\end{itemize}

\section{Upkeep and Maintenance}
\label{subsec:upkeep-options}
\index{upkeep}\index{assets!upkeep}\index{followers!upkeep}

Both assets and followers require regular maintenance. Each downtime period, choose one of the two options.

\subsection*{Asset Upkeep}
\label{sec:asset-upkeep}
Two options:
\begin{itemize}
\item \textbf{Efficient (Higher XP, Less Time)}:  
  Pay Upkeep XP = $\max(1, \lceil \text{Acquisition XP} / 3 \rceil)$, minimal effort.
\item \textbf{Intensive (Lower XP, More Time)}:  
  Pay 1 XP, use a dedicated downtime action with significant personal involvement.
\item \textbf{Failure to Pay}: Asset becomes \emph{Neglected} (or \emph{Compromised} if already Neglected).
\end{itemize}

\subsection*{Follower Upkeep}
\label{sec:follower-upkeep}
Two options:
\begin{itemize}
\item \textbf{Efficient}:  
  Pay Upkeep XP = $\max(1, \lceil \text{Cap}^2 / 3 \rceil)$, minimal effort.
\item \textbf{Intensive}:  
  Pay 1 XP, use a dedicated downtime action with significant personal involvement.
\item \textbf{Failure to Pay}: Follower becomes \emph{Wary} (or \emph{Seized} if already Wary).
\end{itemize}

\subsection*{Upkeep Examples}
\label{sec:upkeep-examples}
\begin{itemize}
  \item \textbf{Minor Asset (4 XP)}: Efficient = 2 XP (4/3 rounded up); Intensive = 1 XP + action.
  \item \textbf{Standard Asset (8 XP)}: Efficient = 3 XP (8/3 = 2.66 → 3); Intensive = 1 XP + action.
  \item \textbf{Cap 3 Follower (9 XP)}: Efficient = 3 XP (9/3); Intensive = 1 XP + action.
  \item \textbf{Cap 2 Follower (4 XP)}: Efficient = 2 XP (4/3 rounded up); Intensive = 1 XP + action.
\end{itemize}

\section{Strategic Considerations}
\label{sec:asset-follower-strategy}
\index{strategic considerations}

\subsection*{When to Invest in Assets}
\label{sec:invest-assets}
\begin{itemize}
\item You need reliable off‑screen capabilities.
\item Your character concept involves wealth or influence.
\item The party lacks certain logistical support.
\item You want to build long‑term influence.
\end{itemize}

\subsection*{When to Invest in Followers}
\label{sec:invest-followers}
\begin{itemize}
\item You need on‑screen assistance.
\item Your character works better with support.
\item The party needs specific capabilities you lack.
\item You want character‑driven story opportunities.
\end{itemize}

\subsection*{Investment by Player Archetype}
\label{sec:invest-by-archetype}
\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Archetype} & \textbf{Typical XP Investment} & \textbf{Best For} \\
\midrule
Solo & 0--10\% & Reliable personal power, minimal upkeep. \\
Mixed (Balanced) & 15--25\% & One solid Follower or Asset (scout, safehouse). \\
Mastermind & 35--55\% & Networks, informants, multiple Assets; build a faction. \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Balance Recommendations}
\label{sec:asset-follower-balance}
\begin{itemize}
\item \textbf{Personal Path} (see \ref{sec:xp-personal-path}): 0--10\% of XP in assets/followers.
\item \textbf{Balanced Path} (see \ref{sec:xp-balanced-path}): 15--25\%.
\item \textbf{Influencer Path} (see \ref{sec:xp-influencer-path}): 35--55\%.
\end{itemize}

\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=Planning for Succession (The Legacy Engine)]
\index{succession}\index{legacy engine}
A loyal Follower (Cap 3+) can become your \textbf{successor} when your character retires or falls -- inheriting your faction standing, one unresolved burden, and a Legacy Bond. See \textbf{Chapter 13: The Legacy Engine} (\ref{ch:legacy}) for full rules. If you plan for your character's story to continue beyond them, invest in a follower you would trust to carry your banner.
\end{tcolorbox}

\section{Loyalty and Relationships}
\label{sec:follower-loyalty}
\index{loyalty}\index{relationships}

\subsection*{Loyalty Levels (Optional)}
\label{sec:loyalty-levels}
\begin{description}
\item[Wary] Cautious, may leave if pressured; +1 XP upkeep cost.
\item[Steady] Reliable, standard performance; normal upkeep.
\item[Devoted] Loyal, may sacrifice; can convert one major complication to a minor setback per arc.
\end{description}

\subsection*{Building Loyalty}
\label{sec:building-loyalty}
\begin{itemize}
\item Fair treatment and respect.
\item Sharing rewards and successes.
\item Protecting followers from harm.
\item Honoring agreements and promises.
\end{itemize}

\subsection*{Losing Loyalty}
\label{sec:losing-loyalty}
\begin{itemize}
\item Mistreatment or disrespect.
\item Unreasonable demands or risks.
\item Broken promises or betrayal.
\item Consistent neglect.
\end{itemize}

\section{Advanced Follower Management}
\label{sec:advanced-followers}
\index{followers!advanced}

\subsection*{Follower Groups}
\label{sec:follower-groups}
For multiple similar followers, you can manage them as a group:
\begin{itemize}
\item \textbf{Single Rating}: Treat as one entity with combined capability.
\item \textbf{Condition Tracking}: Group shares exposure and harm.
\item \textbf{Maintenance}: Single upkeep cost for the group.
\item \textbf{Risks}: Problems affect the entire group.
\end{itemize}

\subsection*{Follower Advancement}
\label{sec:follower-advancement}
Followers can improve over time:
\begin{itemize}
\item \textbf{Experience}: Gain capability through successful assistance.
\item \textbf{Training}: Spend XP to improve follower capabilities.
\item \textbf{Equipment}: Better gear can enhance effectiveness.
\item \textbf{Limits}: Followers typically cap at lower levels than PCs.
\end{itemize}

\section{Risk Management}
\label{sec:asset-follower-risk}
\index{risk management}

\subsection*{Asset Risks}
\label{sec:asset-risks}
\begin{itemize}
\item \textbf{Financial}: Assets can be costly to maintain.
\item \textbf{Security}: Assets can be attacked or stolen.
\item \textbf{Attention}: Valuable assets draw unwanted notice.
\item \textbf{Dependency}: Over‑reliance can be problematic.
\end{itemize}

\subsection*{Follower Risks}
\label{sec:follower-risks-list}
\begin{itemize}
\item \textbf{Safety}: Followers can be harmed or captured.
\item \textbf{Loyalty}: Followers may betray or leave.
\item \textbf{Attention}: Followers can draw enemy interest.
\item \textbf{Morale}: Followers have needs and limits.
\end{itemize}

\subsection*{Mitigation Strategies}
\label{sec:risk-mitigation}
\begin{itemize}
\item \textbf{Diversification}: Don"t put all resources in one place.
\item \textbf{Security}: Protect valuable assets and followers.
\item \textbf{Relationships}: Maintain good terms with your people.
\item \textbf{Contingencies}: Have backup plans for losses.
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title={Assets and Followers Quick Reference},fonttitle=\bfseries]
\textbf{Assets:}
\begin{itemize}
\item Minor: 4 XP \;|\; Standard: 8 XP \;|\; Major: 12 XP
\item Free off‑screen use: once per session
\item Boon activation: spend 1 Boon for scene impact
\item Conditions: \emph{Maintained} $\rightarrow$ \emph{Neglected} $\rightarrow$ \emph{Compromised}
\end{itemize}

\textbf{Followers:}
\begin{itemize}
\item Cost: $\text{Cap}^2$ XP
\item Assistance: $+\min(\text{Cap},\text{skill})$ dice (max +3 from all sources)
\item Independent action: once per scene (party‑wide)
\item Conditions: \emph{Exposure} and \emph{Harm} tracks
\end{itemize}

\textbf{Upkeep Options:}
\begin{itemize}
\item Efficient: $\max(1,\lceil\text{Cost}/3\rceil)$ XP, minimal time
\item Intensive: 1 XP, dedicated downtime action
\end{itemize}
\end{tcolorbox}

\section{Practical Examples}
\label{sec:asset-follower-examples}

\subsection*{Asset Example: The Safehouse}
\label{sec:example-safehouse}
\begin{itemize}
\item \textbf{Type}: Minor Asset (4 XP)
\item \textbf{Free Use}: Secure lodging, basic supplies between adventures.
\item \textbf{Boon Activation}: Reveal a hidden escape route during pursuit.
\item \textbf{Upkeep}: Efficient: 2 XP (4/3 rounded up); Intensive: 1 XP + downtime action.
\item \textbf{Risks}: Discovery by enemies, maintenance costs.
\end{itemize}

\subsection*{Follower Example: The Scout}
\label{sec:example-scout}
\begin{itemize}
\item \textbf{Capability}: 3 (9 XP cost)
\item \textbf{Assistance}: +3 dice on tracking and survival rolls.
\item \textbf{Independent Action}: Scout ahead to improve party position.
\item \textbf{Upkeep}: Efficient: 3 XP (9/3); Intensive: 1 XP + downtime action.
\item \textbf{Risks}: Injury in dangerous scouting; disloyalty if mistreated.
\end{itemize}

\subsection*{Combination Example: The Merchant}
\label{sec:example-merchant}
\begin{itemize}
\item \textbf{Assets}: Trading post (8 XP), caravan (4 XP) --- \emph{12 XP total}
\item \textbf{Followers}: Cap 2 guards (4 XP each = 8 XP), Cap 3 factor (9 XP) --- \emph{17 XP total}
\item \textbf{Total Investment}: \textbf{29 XP} in assets and followers
\item \textbf{Upkeep (Efficient)}: Assets 4 XP + Followers 6 XP = \textbf{10 XP} per downtime
\item \textbf{Benefits}: Trade income, transport, protection, business contacts
\item \textbf{Risks}: Competition, bandit attacks, employee issues, regulatory attention
\end{itemize}

\emph{Remember:} Assets and followers can greatly expand your capabilities, but they require careful management and carry significant risks. Invest wisely based on your character concept and the needs of your group.

\section{Narrative‑Heavy Asset and Follower Options}
\label{sec:narrative-assets}

For groups that prefer strong narrative focus, consider these optional approaches.

\paragraph{Story‑Driven Upkeep}
\label{sec:story-upkeep}
Instead of tracking XP costs for upkeep, the GM can introduce narrative complications that require attention. A neglected asset might attract unwanted attention, while a neglected follower might request a favor or special treatment.

\paragraph{Collaborative Management}
\label{sec:collab-management}
Players can describe how they maintain their assets and followers through roleplay rather than mechanical upkeep costs. A well‑described scene of tending a workshop or bonding with followers can fulfill maintenance requirements.

\paragraph{Asset and Follower as Character Development}
\label{sec:asset-character-dev}
Use asset and follower management as opportunities for character growth and backstory development, allowing players to narrate how their relationships and holdings evolve through significant story moments.

\paragraph{Flexible Condition Tracking}
\label{sec:flexible-conditions}
Focus on the narrative implications of asset and follower conditions rather than strict mechanical penalties. A ``Neglected'' asset might still function but with interesting complications, while a ``Compromised'' asset might require creative solutions rather than just XP investment.

\section{Terrestrial Patrons}
\label{sec:terrestrial-patrons}
\index{Patrons!terrestrial}

Not all power comes from cosmic entities or divine forces. Sometimes your greatest allies---and most demanding creditors---are mortal factions who can offer protection, resources, and influence in the world of men.

\subsection{What Are Terrestrial Patrons?}
\label{sec:terrestrial-definition}

Terrestrial Patrons represent your ongoing relationships with powerful mortal organizations:
\begin{itemize}
    \item Noble houses and courtly factions.
    \item Merchant guilds and trading consortiums.
    \item Military units and mercenary companies.
    \item Criminal organizations and smuggling rings.
    \item Religious orders and temple hierarchies.
    \item Scholarly colleges and arcane societies.
\end{itemize}

Unlike supernatural Patrons (see \ref{ch:magic}), Terrestrial Patrons do not grant magic, but they offer \textbf{leverage}: protection, resources, sanctuary, information, and political influence.

\subsection{Gaining a Terrestrial Patron}
\label{sec:terrestrial-gaining}

To gain a Terrestrial Patron, complete one of these significant actions:
\begin{itemize}
    \item Complete a major job or service for them.
    \item Swear a formal Oath of service.
    \item Enter into legal or financial binding.
    \item Share criminal secrets or be compromised by them.
    \item Perform a notable service that earns their favor.
\end{itemize}
Mark the Patron on your character sheet and write one sentence: \emph{``They want me because \_\_\_\_\_\_\_''}.

\subsection{Obligation System}
\label{sec:terrestrial-obligation}

Terrestrial Patrons use the same Obligation track as supernatural Patrons (see \ref{sec:obligation-capacity}), but consequences are social, legal, or economic rather than mystical.

\begin{itemize}
    \item \textbf{Calling in Favors}: When you request Patron influence, add +1 Obligation.
    \item \textbf{Capacity}: Spirit + Presence (same as supernatural Patrons).
    \item \textbf{Overflow}: Each segment above capacity inflicts 1 Fatigue.
    \item \textbf{Resolution}: Reduce through service, downtime actions, or fulfilling demands.
\end{itemize}

\subsection{Patron Perks}
\label{sec:terrestrial-perks}

Each Terrestrial Patron offers 2--3 repeatable benefits that require no rolls to access.

\textbf{Common Perks Include:}
\begin{itemize}
    \item Sanctuary in their territory or establishments.
    \item Legal relief or protection from certain authorities.
    \item Access to black market goods or specialized services.
    \item Elite followers or hirelings at reduced cost.
    \item Forged documents or identity protection.
    \item Military backing or private security.
    \item Information networks and rumor access.
    \item Trade advantages or exclusive contracts.
\end{itemize}
Using a perk never requires a roll---Fate has already been paid. Each use comes with the ongoing cost of Obligation.

\subsection{Patron Demands}
\label{sec:terrestrial-demands}

Terrestrial Patrons always want something in return:
\begin{itemize}
    \item Silence about sensitive information.
    \item Loyalty over other conflicting interests.
    \item Completion of specific jobs or tasks.
    \item Delivery of valuable names or secrets.
    \item Political or social support for their causes.
    \item Elimination of rivals or threats to their interests.
\end{itemize}
\textbf{Refusing a Demand} raises Obligation by 1 segment and may trigger immediate complications.

\subsection{When Obligation Fills}
\label{sec:terrestrial-obligation-fill}

At \textbf{6 Obligation segments}, the Patron acts---this is not optional. Choose one:
\begin{enumerate}
    \item \textbf{Unavoidable Job}: You must complete a task you cannot refuse.
    \item \textbf{Severe Price}: Pay significant cost (legal, social, or material).
    \item \textbf{First Strike}: They act against you---reputation damage, warrants, bounty, blackmail.
\end{enumerate}
After the consequence lands, reduce Obligation to 3 segments.

\subsection{Cutting Ties}
\label{sec:terrestrial-cutting-ties}

You may sever a Terrestrial Patron relationship, but doing so has serious fallout:
\begin{itemize}
    \item Lose all current perks immediately.
    \item Gain a new \textbf{Rival} faction that opposes your interests.
    \item Take a lasting \textbf{Complication} (Curse, Bounty, or Scandal) that follows you.
    \item May trigger immediate retaliation depending on the Patron's nature.
\end{itemize}
Some Patrons never forgive betrayal; others can be bought off with significant compensation.

\subsection{Redemption and Favor}
\label{sec:terrestrial-redemption}

If you perform a monumental service for a Patron---something beyond what was asked---reduce Obligation by 2 and gain a permanent \textbf{Favor}.

\textbf{Favor Benefits:}
\begin{itemize}
    \item Noble title or official position.
    \item Land grant or property rights.
    \item Permanent access to exclusive resources.
    \item Unique asset or specialized equipment.
    \item Political influence or legal immunity.
    \item Protection from their rivals or enemies.
\end{itemize}

\subsection{Example: The Black Ledger}
\label{sec:terrestrial-example}

The Black Ledger is a smuggling syndicate operating in the coastal city of Kahfagia.

\textbf{Gaining Their Favor:} Successfully complete three smuggling runs without detection.

\textbf{Perks:}
\begin{itemize}
    \item Sanctuary in hidden safehouses throughout the city.
    \item Access to rare imported goods and contraband.
    \item Information about port authorities and customs schedules.
\end{itemize}

\textbf{Demands:}
\begin{itemize}
    \item Maintain silence about syndicate operations.
    \item Transport occasional ``special packages'' (may be illegal).
    \item Eliminate rival smugglers who threaten their routes.
\end{itemize}

\textbf{Sample Scenario:} Rellan calls on the Ledger for a smuggled border crossing. The GM rules it succeeds automatically, but adds +1 Obligation. Later, the Ledger demands he silence a witness who saw too much. If he refuses, Obligation rises again. If Obligation ever reaches 6, the Ledger collects: accounts frozen, bounty posted, or a rival informant sent after him.

\subsection{Creating Your Own Terrestrial Patrons}
\label{sec:terrestrial-creating}

When designing a Terrestrial Patron, consider:
\begin{enumerate}
    \item \textbf{Nature}: What type of organization are they?
    \item \textbf{Motivation}: What drives their goals and actions?
    \item \textbf{Resources}: What can they realistically offer?
    \item \textbf{Methods}: How do they typically operate?
    \item \textbf{Enemies}: Who opposes their interests?
    \item \textbf{Perks}: 2--3 concrete benefits they provide.
    \item \textbf{Demands}: 3--5 typical requests they make.
\end{enumerate}

\emph{Remember:} Terrestrial Patrons make the mortal world feel alive and reactive. They represent the complex web of loyalty, debt, and influence that shapes everyday life in Fate's Edge.

\section{Multi-Character Switching System}
\label{sec:multi-character-switching}
\index{Echo Character}\index{Downtime Log}

This framework allows a player to temporarily move a character offscreen while continuing play with another. The absent character remains active in the fiction and continues shaping the world through \textbf{Downtime Logs}, consequences, and story pressure.

\subsection{Core Principles}
\label{sec:mcs-principles}
\begin{itemize}
    \item Players \textbf{choose} when their character steps offscreen.
    \item Offscreen characters still \textbf{matter} to the narrative.
    \item Absence creates \textbf{opportunity and risk}, not punishment.
    \item The world reacts to characters \textbf{even when unseen}.
\end{itemize}

\subsection{Switch Mechanics}
\label{sec:mcs-switch}

\subsubsection{Declaration}
When a player chooses to step out, they declare:
\begin{itemize}
    \item Which character is going offscreen.
    \item Where the character is going (fictional location).
    \item Why they are stepping out (story reason).
    \item How long they expect to be away (1--3 sessions).
\end{itemize}
The character becomes an \textbf{Echo Character}: present in the world, but offscreen.

\subsubsection{Controlled Status Types}
Echo Characters take one of three statuses:
\begin{itemize}
    \item \textbf{Supporting} (default): Works to assist active PCs.
    \item \textbf{Advancing}: Pursues their own agenda or project.
    \item \textbf{In Peril}: Their absence creates danger or urgency.
\end{itemize}
The GM and player collaborate to set this status.

\subsection{Downtime Log}
\label{sec:mcs-downtime-log}
Each offscreen session, the player completes a simple \textbf{Downtime Log}:

\subsubsection{Structure}
\begin{itemize}
    \item \textbf{Resource Action}: How they maintain connections or standing.
    \item \textbf{Project Action}: One goal or project advanced.
    \item \textbf{Event Response}: How they react to unseen events or threats.
\end{itemize}

\subsubsection{Tier-Based Project Limits}
\begin{itemize}
    \item Tier I--II: 1 Project.
    \item Tier III--IV: 2 Projects.
    \item Tier V: 3 Projects.
\end{itemize}

\subsubsection{GM Response}
The GM records:
\begin{itemize}
    \item The fictional outcome of each declaration.
    \item One meaningful \textbf{Consequence} (complication, shift, or new opportunity).
\end{itemize}

\subsection{Return Mechanics}
\label{sec:mcs-return}

\subsubsection{Re-Entry Process}
When returning to play:
\begin{enumerate}
    \item The returning character gains \textbf{2 Boons} for narrative re‑entry.
    \item Roll \textbf{Wits + Spirit} (DV 3) to reestablish presence.
    \item Address unresolved plot threads created during absence.
\end{enumerate}

\subsubsection{Re-Entry Consequences}
The GM assigns one:
\begin{itemize}
    \item \textbf{Ripple Effect} --- your absence reshaped the situation.
    \item \textbf{Opportunity Missed} --- a door closed without you.
    \item \textbf{Allies Gained} --- your actions forged new connections.
    \item \textbf{Rivals Made} --- someone now works against you.
    \item \textbf{Personal Growth} --- you return wiser or transformed.
    \item \textbf{Lost Momentum} --- your projects slipped.
\end{itemize}

\subsection{Advanced Systems}
\label{sec:mcs-advanced}

\subsubsection{Buddy System}
If multiple players are absent:
\begin{itemize}
    \item Nominate a \textbf{Buddy} to help track Downtime Logs.
    \item The Buddy gains \textbf{+1 additional Boon} for the support role.
\end{itemize}

\subsubsection{Crisis Management}
Offscreen characters track a \textbf{Crisis Clock [6]}:
\begin{itemize}
    \item Fills when negative consequences strike.
    \item At 6: The character faces a major offscreen crisis requiring GM action.
    \item May be reduced through Downtime Log responses or assistance from active PCs.
\end{itemize}

\subsection{GM Guidelines}
\label{sec:mcs-gm}
\begin{itemize}
    \item Spend 5--7 minutes per absent character completing their Downtime Log.
    \item Focus on narrative change, not bookkeeping pressure.
    \item Absence should generate hooks and stakes, not penalties.
    \item Use Echo Characters to keep the world \emph{alive and reactive}.
\end{itemize}

\subsection{Player Guidelines}
\label{sec:mcs-player}
\begin{itemize}
    \item Declare offscreen goals with clear fictional intent.
    \item Treat absence as character growth, not disappearance.
    \item Collaborate with the GM to weave consequences into play.
    \item Expect absence to create tension, opportunity, and drama.
\end{itemize}