\chapter{Compendium of Assets}
łabel{ch:compendium−assets}

∈dex{assets}∈dex{holdings}∈dex{influence}
In \textit{Fate's Edge}, \textbf{Assets} are your off−screen levers−−−properties, charters, networks, and institutions that change the fiction between scenes. They do not fight beside you (see §\ref{ch:assets−followers} for \emph{Followers}) but they shape the ground you walk on.

\section{Design Principles}
∈dex{assets!design principles}
Assets are written to:
\begin{itemize}
  ℹtem \textbf{Shift fiction} between scenes: open doors, quiet rumors, stock larders, calm crowds.
  ℹtem \textbf{Create hooks}: each Asset suggests opportunities \emph{and} complications if neglected.
  ℹtem \textbf{Stay narrative−first}: prefer position, scale, and access over flat modifiers.
\end{itemize}

Assets do \textbf{not}:
\begin{itemize}
  ℹtem Act inside a scene on their own (Followers do that, §\ref{ch:assets−followers}). ∈dex{followers|see{assets and followers}}
  ℹtem Print wealth or XP automatically.
  ℹtem Stack identical tags−−−if two Assets would grant the same edge, take the best one. ∈dex{assets!stacking}
\end{itemize}

\paragraph{Narrative First, Mechanics Second}
When an Asset grants ``start \textbf{Dominant}'', it applies to social or informational rolls where positioning matters. For purely practical challenges (e.g., picking a lock, repairing a device), ``reduce DV by 1'' is the preferred phrasing. When in doubt, the GM chooses the form that best fits the fiction.

\section{When to Invest in Assets}
∈dex{assets!investment advice}
The table below matches player archetypes (see §\ref{sec:player−archetypes}) to typical Asset investment. Use it as a guideline, not a straitjacket.

\begin{center}
\begin{tabular}{lll}
⊤rule
\textbf{Archetype} & \textbf{Typical XP in Assets} & \textbf{Recommended Assets} \\
∣rule
Solo & 0−−10% & None or one Minor (e.g., Signature Weapon, Scholar's Cell) \\
Mixed (Balanced) & 15−−░ & One or two Standard Assets (e.g., Safehouse, Informant Ring) \\
Mastermind & 35−−55% & Multiple Standard or Major Assets; aim to control a faction (see Terrestrial Patrons, §\ref{ch:world−powers−obligation}) \\
⊥tomrule
\end{tabular}
\end{center}

\paragraph{Succession Note}
A loyal follower (Cap 3+) can inherit your faction or network, but Assets themselves are not directly inheritable unless you have converted a terrestrial patron into a personal Asset (see §\ref{sec:takeover−sidebar}). For a character's death or retirement, the GM may allow a \textbf{successor} (see \ref{ch:legacy−engine}) to retain one Standard Asset as a ``legacy'' with suitable narrative justification (e.g., the family safehouse, the guild seat passes to the heir).

\section{Asset Mechanics (Summary)}
∈dex{assets!mechanics}∈dex{upkeep}∈dex{Story Beats (SB)}
All Assets share the following properties (expanded rules: §\ref{ch:assets−followers}):
\begin{description}
  ℹtem[Cost] Minor \textbf{4 XP}, Standard \textbf{8 XP}, Major \textbf{12 XP}. ∈dex{assets!minor}∈dex{assets!standard}∈dex{assets!major}
  ℹtem[Establishment] 1 day / 1 week / 1 month respectively to become operational.
  ℹtem[Upkeep] Each downtime, either pay XP equal to the Asset's base cost \emph{or} devote a full scene tending it. Neglect degrades condition (Maintained $→$ Neglected $→$ Compromised). ∈dex{assets!conditions}∈dex{assets!maintenance}
  ℹtem[Free Use] Once per session, trigger the listed \emph{off−screen} benefit (framed at scene start or as flashback with GM assent).
  ℹtem[Scene Surge] Spend \textbf{1 Boon} to turn an Asset into a dramatic \emph{on−screen} intervention as described (one beat, fiction permitting). ∈dex{Boons}
  ℹtem[Extra Use] Spend \textbf{2 XP} to trigger the off−screen benefit an additional time this session.
  ℹtem[Risk Interface] When an Asset covers risk, it typically \textbf{improves Position}, \textbf{lowers DV by 1}, \textbf{fills 1 segment on a relevant timer}, or \textbf{mitigates 1 SB} (Story Beat) from a thematically linked complication. ∈dex{position}∈dex{difficulty value (DV)}∈dex{timers}∈dex{Story Beats (SB)}
\end{description}

\paragraph{Capacity (Optional Rule)}∈dex{assets!capacity}
A character comfortably maintains a number of Assets up to their \textbf{Spirit} rating. For each Asset beyond this, increase its upkeep by +▒ (round up) \emph{or} require an additional downtime scene to keep it \emph{Maintained}.

\emph{Example: Spirit 3 $→$ 3 Assets are easy; a 4th Asset costs ▒ more upkeep (e.g., 6 XP instead of 4) each downtime, or demands a second personal scene to maintain.}

\paragraph{Cross‑Reference to Faction Ties}
Assets tied to a specific faction (e.g., Temple Complex, Banking Charter, Guild Seat) may also generate \textbf{Obligation} if used recklessly or ignored. See §\ref{ch:world−powers−obligation} for terrestrial Patron mechanics.

\section{Minor Assets (4 XP)}
∈dex{assets!minor}
Small holdings, charters, or kits that grant reliable leverage.
\begin{description}
  ℹtem[\textbf{Signature Weapon}] ∈dex{signature weapon} \emph{A named heirloom or masterpiece.} \emph{Free Use:} If gear damage would compromise it, downgrade to a minor flaw and clear with a quiet maintenance scene. \emph{Surge:} Produce it at the perfect moment; improve Position one step for your first attack this scene.

  ℹtem[\textbf{Scholar's Cell}] ∈dex{scholar's cell} \emph{A quiet room, permits, and a custodian.} \emph{Free Use:} Begin one research scene at \textbf{Dominant} (or lower DV by 1). \emph{Surge:} Reveal a cited source now; fill 1 segment on a relevant \emph{Information} timer.

  ℹtem[\textbf{Safehouse Network (Quarter)}] ∈dex{safehouse} \emph{Trapdoors, false walls, and a landlord who never saw you.} \emph{Free Use:} Start an entry/exit scene \textbf{Dominant}. \emph{Surge:} Produce a hidden egress; convert one \emph{pursuit} consequence into a temporary complication you can interact with.

  ℹtem[\textbf{Artisan's Bench}] ∈dex{artisan's bench} \emph{Tools for field repairs and fabrication.} \emph{Free Use:} Remove \emph{Compromised} from one mundane item between scenes. \emph{Surge:} Improvise a single−use tool; gain a \emph{Boon} on the immediate Craft check.

  ℹtem[\textbf{Courier Chain}] ∈dex{courier network} \emph{Runners, dead−drops, and codes in one city.} \emph{Free Use:} Move a message or small parcel safely; start the related parley \textbf{Dominant}. \emph{Surge:} A courier appears now; mitigate 1 SB tied to a missed appointment or lost message.

  ℹtem[\textbf{Small Holding (Farm/Orchard)}] ∈dex{farm}∈dex{orchard} \emph{Steady larder and a trusted foreman.} \emph{Free Use:} Waive a minor Supply complication or clear Fatigue 1 for one PC during downtime. \emph{Surge:} Produce a timely wagon, mule, or crates to justify a plausible cover story.

  ℹtem[\textbf{Neighborhood Inn/Tavern}] ∈dex{inn}∈dex{tavern} \emph{A familiar house of bread and rumor.} \emph{Free Use:} First social opener in that neighborhood starts \textbf{Dominant}. \emph{Surge:} Call in the proprietor; shift a hostile crowd one step toward neutral.

  ℹtem[\textbf{Ritual Niche/Shrine}] ∈dex{shrine}∈dex{ritual niche} \emph{A consecrated corner with a small flame.} \emph{Free Use:} Clear Fatigue 1 or mitigate 1 SB from spiritual or psychological complications (fear, despair, doubt). \emph{Surge:} Sanctify a doorway for one scene; hostile supernatural presence must reveal itself or back away one band (§\ref{ch:world−interaction}).

  ℹtem[\textbf{River Skiff/Barge Share}] ∈dex{boat}∈dex{barge} \emph{Limited access to a hull and a taciturn pilot.} \emph{Free Use:} Water crossing scenes begin \textbf{Dominant}. \emph{Surge:} Evade a patrol with a practiced channel; reduce DV by 1 on the escape roll.

  ℹtem[\textbf{Herbal Plot}] ∈dex{herbs}∈dex{garden} \emph{Drying racks, clay pots, and a steady hand at dawn.} \emph{Free Use:} Remove Fatigue 1 from one character or gain a \emph{Boon} to a single mundane healing attempt this downtime. \emph{Surge:} Produce a rare poultice; convert a Severe healing timer into two Moderate ticks.

  ℹtem[\textbf{Personal Library}] ∈dex{library!personal} \emph{Shelves of annotated tomes and a reading chair.} \emph{Free Use:} Gain a \emph{Boon} on one Lore roll per session. \emph{Surge:} Quote a relevant passage; reduce DV of a related Investigation by 1.

  ℹtem[\textbf{Stable/Coach House}] ∈dex{stable}∈dex{transport} \emph{Reliable mounts and drivers who ask few questions.} \emph{Free Use:} Land travel scenes begin \textbf{Dominant}. \emph{Surge:} Produce a fresh mount or spare vehicle; ignore the first \emph{exhaustion} complication.

  ℹtem[\textbf{Alchemist's Workshop}] ∈dex{alchemist} \emph{Reagents, retorts, and a suspicious chimney.} \emph{Free Use:} Produce one dose of a common alchemical item between scenes. \emph{Surge:} Brew a specialized compound; gain a \emph{Boon} on the next Craft (Alchemy) roll.

  ℹtem[\textbf{Messenger Birds}] ∈dex{messenger birds}∈dex{communication} \emph{Trained pigeons or ravens with coded leg bands.} \emph{Free Use:} Send a simple message across moderate distance safely. \emph{Surge:} Receive urgent news now; mitigate 1 SB from \emph{miscommunication}.
\end{description}

\section{Standard Assets (8 XP)}
∈dex{assets!standard}
Established holdings with regional reach.
\begin{description}
  ℹtem[\textbf{Charter or Minor Title}] ∈dex{charter}∈dex{title} \emph{Recognized papers and privileges.} \emph{Free Use:} Secure an audience or official hearing; start \textbf{Dominant}. \emph{Surge:} Invoke prerogative to delay immediate sanctions; mitigate 1 SB of civic backlash.

  ℹtem[\textbf{Guild Seat}] ∈dex{guild} \emph{Voting rights and shop‑rights.} \emph{Free Use:} Commission or procurement opens \textbf{Dominant} or DV $−1$. \emph{Surge:} A guild factor vouches; fill 1 segment on a \emph{Permit/Access} timer.

  ℹtem[\textbf{Market Front}] ∈dex{market}∈dex{stall} \emph{A fixed stall and regular customers.} \emph{Free Use:} Treat one \emph{Minor} Asset's upkeep as waived this downtime. \emph{Surge:} Liquidate stock; justify on−hand mundane gear within reason.

  ℹtem[\textbf{Informant Ring}] ∈dex{spies}∈dex{informants} \emph{Eyes and ears on retainer.} \emph{Free Use:} Targeted inquiry begins \textbf{Dominant}. \emph{Surge:} Reveal a hidden schedule or route now; mitigate 1 SB from \emph{ambush/surprise}.

  ℹtem[\textbf{Mercenary Contract (Cap 2)}] ∈dex{mercenaries} \emph{Access to a small trained unit.} \emph{Free Use:} Introduce temporary \emph{off−screen} security that downgrades ``raid'' to ``attempted raid.'' \emph{Surge:} One on−screen intervention (one beat) that improves Position for a withdrawal or breach.

  ℹtem[\textbf{Trading Charter}] ∈dex{trade}∈dex{charter!trading} \emph{Right−of−way at gates and tolls.} \emph{Free Use:} Border/inspection scenes start \textbf{Dominant}. \emph{Surge:} Produce stamped manifests; DV $−1$ on the next scrutiny roll.

  ℹtem[\textbf{Licensed Library}] ∈dex{library} \emph{Stacks and a learned custodian.} \emph{Free Use:} Deep‑topic research begins \textbf{Dominant} and gains a \emph{Boon}. \emph{Surge:} Cite a precedent to force a pause in debate (one exchange).

  ℹtem[\textbf{Craftsman's Quarter}] ∈dex{workshops}∈dex{apprentices} \emph{Shop plus apprentices.} \emph{Free Use:} Clear \emph{Compromised} from a crafted item or structure. \emph{Surge:} Rush a bespoke piece; fill 1 segment on a \emph{Build/Repair} timer.

  ℹtem[\textbf{Temple Complex}] ∈dex{temple} \emph{Courts, choirs, charity.} \emph{Free Use:} Among adherents, start \textbf{Dominant} or mitigate 1 SB from \emph{crowd/morale}. \emph{Surge:} Sanctuary: hostile social pressure must escalate openly or wait.

  ℹtem[\textbf{Harbor Office}] ∈dex{harbor}∈dex{port} \emph{Berths, booms, bells.} \emph{Free Use:} Priority loading/unloading; start \textbf{Dominant}. \emph{Surge:} Hold the chain for one beat; negate a single \emph{pursuit by water} consequence.

  ℹtem[\textbf{Healing House}] ∈dex{healing}∈dex{infirmary} \emph{Beds, herbs, bells.} \emph{Free Use:} During downtime, clear \emph{Harm 1} or \emph{Fatigue 2} from one ally. \emph{Surge:} Stabilize now; convert a Severe injury consequence into a 4‑segment Recovery timer.

  ℹtem[\textbf{Observatory}] ∈dex{observatory}∈dex{navigation} \emph{Lenses and ephemerides.} \emph{Free Use:} Predict timing; navigation/timing openers start \textbf{Dominant}. \emph{Surge:} Name a narrow window; DV $−1$ for an action attempted inside it.

  ℹtem[\textbf{Performance Venue}] ∈dex{theatre}∈dex{performance} \emph{Stage and regulars.} \emph{Free Use:} Begin Performance with a receptive crowd. \emph{Surge:} Spotlight an ally; shift a crowd one step toward your desired mood.

  ℹtem[\textbf{Hospitality House}] ∈dex{hospitality}∈dex{guest right} \emph{Beds, bread, rumor.} \emph{Free Use:} Gain a \emph{Protected pause} in parley on your ground. \emph{Surge:} Invoke guest−right; hostile social pressure must step into the open.

  ℹtem[\textbf{Storage Vault}] ∈dex{vault}∈dex{storage} \emph{Locks and ledgers, damp stone and iron.} \emph{Free Use (per session):} Ignore the first \emph{theft/sabotage} targeting stored goods. \emph{Surge:} Produce a cached item plausibly stored here.

  ℹtem[\textbf{Hidden Cache}] ∈dex{cache}∈dex{emergency} \emph{A buried lockbox or hollow tree, accessible anywhere in the city.} \emph{Free Use:} Once per session, access emergency gear when cornered; start the scene \textbf{Dominant}. \emph{Surge:} Reveal a larger stash; fill 1 segment on a \emph{Supply} timer.

  ℹtem[\textbf{Signal Relay}] ∈dex{relay}∈dex{communications} \emph{Fires, flags, drums.} \emph{Free Use:} Coordinated actions across distance start \textbf{Dominant}. \emph{Surge:} Push a ready signal now; mitigate 1 SB from \emph{confusion/lag}.

  ℹtem[\textbf{Merchant Caravan}] ∈dex{caravan}∈dex{trade} \emph{One wagon, two guards, and a route.} \emph{Free Use:} Trade missions begin \textbf{Dominant}. \emph{Surge:} Call in the caravan early; provide unexpected supplies or reinforcements.

  ℹtem[\textbf{Legal Practice}] ∈dex{law}∈dex{barrister} \emph{Papers, precedents, and clients.} \emph{Free Use:} Legal proceedings start \textbf{Dominant} or DV $−1$. \emph{Surge:} Produce a forgotten writ; negate one \emph{legal} complication.

  ℹtem[\textbf{Smuggler's Network}] ∈dex{smugglers}∈dex{underground} \emph{Contacts and hidden routes.} \emph{Free Use:} Illicit transport begins \textbf{Dominant}. \emph{Surge:} Bypass a checkpoint; mitigate 1 SB from \emph{customs/patrols}.

  ℹtem[\textbf{Crafting Guildhall}] ∈dex{guildhall}∈dex{crafting} \emph{Tools, masters, and apprentices.} \emph{Free Use:} Complex crafting projects gain a \emph{Boon}. \emph{Surge:} Rush production; fill 1 segment on a \emph{Manufacturing} timer.
\end{description}

\section{Major Assets (12 XP)}
∈dex{assets!major}
Significant holdings that bend regional outcomes.
\begin{description}
  ℹtem[\textbf{Ward License}] ∈dex{city ward}∈dex{license} \emph{You run a city district's levers.} \emph{Free Use:} In−ward civic openers start \textbf{Dominant}. \emph{Surge:} Re−route patrols or light−−−shift a chase route 1 step in your favor.

  ℹtem[\textbf{Exclusive Right (Monopoly)}] ∈dex{monopoly} \emph{Control of a staple (salt, timber, ferries).} \emph{Free Use:} In market scenes, ignore the first \emph{supply} complication. \emph{Surge:} Declare a brief scarcity/surplus to leverage terms; DV $−1$ on a single negotiation.

  ℹtem[\textbf{Fortress/Stronghold}] ∈dex{fortress}∈dex{stronghold} \emph{Stone, stores, signals.} \emph{Free Use:} In raid/siege framing, negate the first \emph{breach} consequence. \emph{Surge:} Sound beacons; allies arrive in time to improve Position for withdrawal or stand.

  ℹtem[\textbf{Demesne (Estate)}] ∈dex{estate}∈dex{demesne} \emph{Fields, mills, tenants.} \emph{Free Use:} Waive minor Supply fallout this arc or treat one additional Minor Asset as having free upkeep. \emph{Surge:} Muster teamsters and hands; fill 1 segment on a \emph{Logistics} timer.

  ℹtem[\textbf{Fleet Share}] ∈dex{fleet} \emph{A cut of hulls and signals.} \emph{Free Use:} Maritime travel starts \textbf{Dominant}. \emph{Surge:} Signal a safe corridor; mitigate 1 SB from \emph{storm/patrol}.

  ℹtem[\textbf{College (University Seat)}] ∈dex{university}∈dex{college} \emph{Faculty, quad, seal.} \emph{Free Use:} In downtime, reduce \emph{one} skill advancement cost by 1 XP (minimum 1). \emph{Surge:} Convene a colloquy to force a pause in magical proceedings (§\ref{ch:magic}).

  ℹtem[\textbf{Banking Charter}] ∈dex{bank}∈dex{letters of credit} \emph{Letters, liens, a counting house.} \emph{Free Use:} Present credit to bypass petty fees and minor delays. \emph{Surge:} Freeze a rival's access for one exchange; mitigate 1 SB from \emph{bribery/extortion}.

  ℹtem[\textbf{Diplomatic Enclave}] ∈dex{diplomacy}∈dex{enclave} \emph{Flag, immunity, ink.} \emph{Free Use:} In foreign courts, begin \textbf{Dominant}. \emph{Surge:} Invoke safe−conduct; hostile action must escalate openly or withdraw.

  ℹtem[\textbf{Mine/Quarry}] ∈dex{mine}∈dex{quarry} \emph{Ore, stone, skilled crews.} \emph{Free Use:} Clear \emph{Compromised} on one structure or armament. \emph{Surge:} Deliver timbers/stone now; DV $−1$ on a \emph{reinforce/repair} roll.

  ℹtem[\textbf{Mining Operation}] ∈dex{mining}∈dex{resources} \emph{Shafts, carts, and crews.} \emph{Free Use:} Ignore the first \emph{resource shortage} complication this arc. \emph{Surge:} Strike a rich vein; produce unexpected wealth or materials.

  ℹtem[\textbf{Military Outpost}] ∈dex{military}∈dex{outpost} \emph{Barracks, stores, and command.} \emph{Free Use:} Military operations begin \textbf{Dominant}. \emph{Surge:} Call for reinforcements; improve Position for a tactical withdrawal or assault.

  ℹtem[\textbf{Trading Consortium}] ∈dex{consortium}∈dex{commerce} \emph{Multiple fronts and shared capital.} \emph{Free Use:} Ignore the first price gouging or supply shortage complication this arc. \emph{Surge:} Coordinate a multi−front deal; DV $−1$ on a major commercial negotiation.

  ℹtem[\textbf{Religious Order}] ∈dex{religious order}∈dex{temple} \emph{Chapters, rites, and influence.} \emph{Free Use:} Among believers, start \textbf{Dominant} or mitigate 1 SB from \emph{heresy/doubt}. \emph{Surge:} Call a council; force a pause in hostile religious proceedings.
\end{description}

\section{Specialized Assets}
∈dex{assets!specialized}
Narrow tools with sharp edges.
\begin{description}
  ℹtem[\textbf{Magical Laboratory} (10 XP)] ∈dex{laboratory!magical}∈dex{rituals} \emph{Wards, benches, focus crystals.} \emph{Free Use:} Once per session, reduce Ritual Backlash by 1 SB (§\ref{ch:magic}). \emph{Surge:} Contain an emergent effect for one beat; begin the control roll \textbf{Dominant}.

  ℹtem[\textbf{Training Academy} (12 XP)] ∈dex{academy}∈dex{training} \emph{Masters and mats.} \emph{Free Use:} In downtime, reduce \emph{one} skill advancement cost by 1 XP (min 1) for a designated trainee. \emph{Surge:} Call in a drilled team to assist off−screen; fill 1 segment on a \emph{Preparation} timer.

  ℹtem[\textbf{Intelligence Bureau} (10 XP)] ∈dex{intelligence}∈dex{analysis} \emph{Analysts and cut−outs.} \emph{Free Use:} Avoid surprise; the first recon opener starts \textbf{Dominant}. \emph{Surge:} Burn a channel; mitigate 1 SB from \emph{counterintelligence}.
\end{description}

\section{Tier III Institutions (16−−20 XP)}
∈dex{assets!tier III}∈dex{institutions}
Empire−shaping engines. Each requires ongoing political play.
\begin{description}
  ℹtem[\textbf{Grand Forge−Citadel} (18 XP)] ∈dex{forge}∈dex{citadel} \emph{City heat and hammer.} \emph{Free Use:} Treat a batch of masterwork gear as immune to the first \emph{gear} complication this arc. \emph{Surge:} Rush a critical piece; fill 2 segments on a \emph{Build/Arm} timer.

  ℹtem[\textbf{Great Library} (16 XP)] ∈dex{great library} \emph{Memory palace of the region.} \emph{Free Use:} Major inquiry begins \textbf{Dominant} and gains a \emph{Boon}. \emph{Surge:} Produce a sealed citation; negate a single \emph{misinterpretation} complication.

  ℹtem[\textbf{Grand Exchange} (20 XP)] ∈dex{exchange}∈dex{bourse} \emph{Markets move when you breathe.} \emph{Free Use:} In high finance, ignore the first \emph{market shock}. \emph{Surge:} Halt trading bells for one exchange; DV $−1$ on one decisive deal.

  ℹtem[\textbf{Fleet Command} (18 XP)] ∈dex{admiralty}∈dex{naval command} \emph{Signals across sea lanes.} \emph{Free Use:} Declare blockade/corridor; hostile egress gains $+1$ step of difficulty. \emph{Surge:} Redirect a squadron; improve Position for a maritime raid or escape.

  ℹtem[\textbf{Hydraulic Works} (16 XP)] ∈dex{aqueduct}∈dex{irrigation} \emph{Water is wealth.} \emph{Free Use:} Cancel the first \emph{shortage/drought} fallout this arc. \emph{Surge:} Flood or feed a field at the critical moment; fill 1 segment on a \emph{Famine Relief} timer.

  ℹtem[\textbf{Palace Complex} (20 XP)] ∈dex{palace}∈dex{court} \emph{Keys to office and omen.} \emph{Free Use:} Political openers start \textbf{Dominant}. \emph{Surge:} Convene a chamber; convert one \emph{scandal} consequence into a public inquiry you can contest.
\end{description}

\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title=Legacy Assets (25+ XP)]
Entire regions, autonomous armies, or a kingdom's treasury. These are campaign anchors, not purchases. Acquire them through story, not XP alone. Examples:
\begin{itemize}
  ℹtem \textbf{March Lordship} −− Rule a border territory; must defend it from raids.
  ℹtem \textbf{Fleet of the Inner Sea} −− Control over a major naval force; attracts enemies and admirals.
  ℹtem \textbf{Imperial Restoration Charter} −− You speak with the voice of a fallen empire, but rivals will never stop scheming.
\end{itemize}
\end{tcolorbox}

\section{Upgrades & Variants}
∈dex{assets!upgrades}∈dex{assets!variants}
Bolster an existing Asset with additional investment (costs stack).
\begin{description}
  ℹtem[\textbf{Enhanced Security} (+2 XP)] ∈dex{security} Hardened doors, sober watchers. \emph{Effect:} Each session, ignore or mitigate 1 SB from theft or sabotage against this Asset.

  ℹtem[\textbf{Expanded Scope} (+3 XP)] ∈dex{scope} Wider footprint or second site. \emph{Effect:} Use the Asset's \emph{Free Use} twice per session \emph{or} across two adjacent districts/routes.

  ℹtem[\textbf{Arcane Wards} (+4 XP)] ∈dex{wards}∈dex{magic!wards} Sanctified and sigiled. \emph{Effect:} Once per session, reduce magical Backlash by 1 SB on−site \emph{or} treat supernatural intrusion as one step easier to detect.

  ℹtem[\textbf{Political Patrons} (+3 XP)] ∈dex{patrons}∈dex{politics} Seals and signatures. \emph{Effect:} In civic scenes tied to this Asset, begin \textbf{Dominant} or DV $−1$.
\end{description}

\section{Creating Your Own Asset}
łabel{sec:creating−your−own−asset}
∈dex{assets!creation}
When drafting a new Asset, answer:
\begin{enumerate}
  ℹtem \textbf{Scope}: Who benefits? (you / party / named faction)
  ℹtem \textbf{Fictional Edge}: What doors open or dangers dim?
  ℹtem \textbf{Cost Tier}: Minor 4 / Standard 8 / Major 12 / Tier III 16−−20.
  ℹtem \textbf{Free Use}: One clear, repeatable off−screen benefit.
  ℹtem \textbf{Scene Surge}: One flashy, plausible \emph{on−screen} move (1 beat) at the cost of 1 Boon.
  ℹtem \textbf{Upkeep Hook}: What happens if you neglect it?
\end{enumerate}

\section{Using Assets in Play}
∈dex{assets!in play}∈dex{downtime}∈dex{flashbacks}
\begin{itemize}
  ℹtem \textbf{Frame the benefit} at scene start (\emph{why is today easier?}).
  ℹtem \textbf{Flash back} with a modest cost to show the Asset's prior work.
  ℹtem \textbf{Tie complications} to Asset condition−−−if it's \emph{Neglected}, the GM has permission to spend SB against it first.
  ℹtem \textbf{Spotlight spread}: Rotate which Assets take the stage so each investment breathes on−screen now and then.
\end{itemize}

\paragraph{Example of Play}
Rellan uses his \textbf{Harbor Office (Standard Asset)} free use to start a customs inspection at \textbf{Dominant} position. Later, when a rival tries to trap him on the wharf, he spends a Boon for a \textbf{Scene Surge}: the harbor chain is raised, blocking pursuit for one beat−−−just enough to slip into a waiting skiff.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Asset Quick Reference,fonttitle=\bfseries]
\begin{itemize}
  ℹtem \textbf{Costs}: Minor 4 XP / Standard 8 XP / Major 12 XP / Tier III 16−−20 XP.
  ℹtem \textbf{Free Use}: Once per session, off−screen leverage (position, DV $−1$, fill 1 timer segment, or mitigate 1 SB).
  ℹtem \textbf{Scene Surge}: Spend 1 Boon for a one−beat, on−screen intervention.
  ℹtem \textbf{Upkeep}: Pay XP = base cost \emph{or} devote a downtime scene; neglect degrades condition.
  ℹtem \textbf{Capacity (Optional)}: Assets $≤$ \textbf{Spirit} are easy; beyond that, upkeep taxes mount.
\end{itemize}
\end{tcolorbox}

\section{Summary}
∈dex{assets!summary}
Minor Assets grease doors; Standard ones bend routes; Major holdings set tables; Tier III institutions tilt the map. Choose what to own, then let the world answer.

\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title=From Faction to Asset: Taking Over a Terrestrial Patron,label=sec:takeover−sidebar]
∈dex{terrestrial patrons!takeover}∈dex{assets!from factions}
Sometimes you stop owing the guild −− and start owning it. Rising to lead a faction (Thieves' Guild, Mercenary Company, Banking Charter, etc.) is a major story arc, not a simple XP purchase. This sidebar outlines how to turn a terrestrial patron into a personal Asset.

\paragraph{Narrative Prerequisites}
Before you can attempt a takeover, you must:
\begin{itemize}
  ℹtem Have at least \textbf{6 Obligation} with the faction (you are deeply entangled).
  ℹtem Complete a \textbf{major service} for the faction (GM's discretion −− e.g., saving it from destruction, winning a war, exposing a traitor at great cost).
  ℹtem Secure at least one \textbf{internal ally} with significant influence (another captain, a guildmaster, a wealthy backer).
\end{itemize}

\paragraph{The Takeover Arc}
The GM and player collaborate on a 2−−3 session arc that typically involves:
\begin{enumerate}
  ℹtem \textbf{Crisis or Power Vacuum} −− The current leader dies, disappears, or is disgraced.
  ℹtem \textbf{Claim of Right} −− You present a valid claim (charter, election, duel victory, popular support).
  ℹtem \textbf{Consolidation} −− Defeat or co‑opt rivals, bribe or intimidate dissenters, and secure the faction's resources.
\end{enumerate}

\paragraph{Mechanical Cost}
Once the arc is complete, you must spend \textbf{12 XP} (the cost of a Major Asset) to convert the faction into a personal Asset. This represents the time, bribes, and institutional restructuring required to make it \emph{yours} rather than a temporary alliance.

\paragraph{Converting the Asset}
The new Asset uses the statistics of the most appropriate existing Asset from this compendium, modified by the faction's nature:
\begin{itemize}
  ℹtem \textbf{Thieves' Guild} $→$ \textbf{Informant Ring} (Standard) or \textbf{Smuggler's Network} (Standard)
  ℹtem \textbf{Mercenary Company} $→$ \textbf{Mercenary Contract} (Minor/Standard) or \textbf{Military Outpost} (Major)
  ℹtem \textbf{Banking Charter} $→$ \textbf{Banking Charter} (Major) or \textbf{Trading Consortium} (Major)
  ℹtem \textbf{Royalist Network} $→$ \textbf{Intelligence Bureau} (Specialized) or \textbf{Diplomatic Enclave} (Major)
  ℹtem \textbf{Ecktorian Mystery Cult} $→$ \textbf{Temple Complex} (Standard) or \textbf{Magical Laboratory} (Specialized)
\end{itemize}
The GM may also create a \textbf{custom Asset} using the guidelines in §\ref{sec:creating−your−own−asset}.

\paragraph{What You Gain}
\begin{itemize}
  ℹtem The Asset's normal \textbf{Free Use} and \textbf{Scene Surge} benefits.
  ℹtem The \textbf{Access Tag} associated with the faction (e.g., \textit{Guild−Ledgered}, \textit{Harbor Handsigns}).
  ℹtem A permanent \textbf{obligation track} toward the faction's members (now your underlings). This track is smaller (4 segments) and fills when you neglect their welfare or make reckless decisions that endanger them.
\end{itemize}

\paragraph{What You Lose}
\begin{itemize}
  ℹtem You can no longer use the faction as a \emph{terrestrial patron} to gain Obligation‑free perks −− you are now responsible for it.
  ℹtem The faction becomes a target for your enemies. The GM may advance a \textbf{Rival Faction timer} whenever you use the Asset aggressively.
  ℹtem You must pay normal \textbf{upkeep} (XP or downtime scenes) to keep the Asset \emph{Maintained}. Neglect causes internal dissent, corruption, or defections.
\end{itemize}

\paragraph{Optional: Gradual Takeover (No XP Cost)}
For a slower, more story‑driven approach, the player can forgo the 12 XP cost but must instead \textbf{spend a downtime scene per session} managing the faction (equivalent to Intensive upkeep) and accept that the Asset's benefits are unreliable until they have successfully completed a full arc of consolidation. The GM may allow the XP cost to be paid in installments (e.g., 4 XP per downtime period) to represent gradual investment.

\paragraph{Example: Rellan Takes the Black Ledger}
Rellan has served the Black Ledger smuggling syndicate for two arcs, accruing 6 Obligation. After exposing a rival who betrayed the syndicate, the aging leader steps down. Rellan spends 12 XP, completes a short arc consolidating loyal captains, and converts the syndicate into a \textbf{Smuggler's Network (Standard Asset)}. He now commands hidden routes and safehouses −− but must keep the crew paid and protected, or an internal revolt timer will start ticking.
\end{tcolorbox}