\chapter{Character Backgrounds}
\label{ch:backgrounds}

% === Epigraph / Intro vignette ===
\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title={From the journal of a Thepyrgosi Witch Hunter},fonttitle=\bfseries]
They call it mercy when the Tower does it and cruelty when I do. The difference is a seal of wax and a hymn at the gate.

I was a bright spark once—could taste iron on my tongue when the air went thin before a working. The masters in Thepyrgos tested me, weighed me, and found me \emph{nearly} enough. ``Nearly'' is a door that never opens again.

So I learned the other doors.

I keep a ledger: names, places, the hour the lamps guttered in a tenement, the way the brick sweats when a child hums without meaning to. I carry writs that make people look away. I carry cord and a promise I do not speak aloud. The University pays for talent delivered in time; the city pays for fires that never start. Between those accounts, I balance what I can.

You will say I go too far. Sometimes I do. Sometimes the street is a tinderbox and a wrong whisper is a spark. Sometimes a family would rather believe in saints than in ash. I give them a story they can live with and take away the one that might kill them.

When the bells of Thepyrgos answer the river wind, I walk the colonnade to the bursar's door and do not look at the windows. The work is ugly. The work is necessary. The night does not care which of those is true—only that the lamps keep burning.
\end{tcolorbox}

Before your character took their first deliberate risk, they belonged somewhere. Backgrounds are the trail-dust on your boots, the school-ink on your fingers, the bell-pattern you still tap when frightened. They explain how you know what you know—and who still knows \emph{you}.

\section{Understanding Backgrounds}
\index{backgrounds}\index{character creation!backgrounds}

Backgrounds serve several important functions:
\begin{itemize}
\item Provide narrative context for abilities and choices
\item Explain how you gained skills, languages (see Chapter \ref{ch:languages} in the full rules), and connections
\item Create ties to regions and factions (\S\ref{ch:world-regions})
\item Offer roleplaying hooks, responsibilities, and privileges
\item Seed reliable \emph{contacts}, \emph{customs}, and \emph{places} that matter on-screen
\end{itemize}

\paragraph{Mechanical Hooks.}
\index{backgrounds!mechanics}
Choose \textbf{one background} at character creation. It grants:
\begin{itemize}
\item \textbf{Access Tags} (narrative permissions; see below)
\item \textbf{A Signature Contact} (named NPC; usually a Cap~1 follower you \emph{do not} pay XP for; they cannot take independent actions (see \S\ref{sec:follower-independent}), but can grant +1d \emph{once per scene} when their help is plausible)
\item \textbf{A Background Boon} (once per session, gain a +1d or reduce DV by 1 for a task strongly tied to the background's sphere)
\item \textbf{An Obligation Clock} (4 segments; when filled, it creates a \emph{call-in} from that world—a favor owed, a duty recalled, a price to pay; see \S\ref{sec:bg-obligation})
\end{itemize}

\paragraph{Access Tags.}
\index{access tags}\index{backgrounds!access}
Tags describe \emph{where} and \emph{with whom} you have standing. Examples: \textit{Guild-Ledgered}, \textit{Court-Announced}, \textit{Veteran-of-the-Marches}, \textit{Bell-Kin of the Causeways}, \textit{Wind-Rider's Tokens}, \textit{Charter-Stamped}. When a scene engages a tag, the GM may:
\begin{itemize}
\item Waive a trivial gate (no roll) or
\item Lower Position risk by one step for first contact or
\item Offer a \emph{Devil's Bargain} tied to that society's expectations
\end{itemize}

\begin{tcolorbox}[colback=black!3!white,colframe=black!40!white,title=Backgrounds and the Legacy Engine]
\index{backgrounds!legacy}\index{legacy engine}
Your \textbf{Signature Contact} is a loyal NPC (Cap 1). Over time, you can upgrade them to a full Follower (pay Cap² XP). Such a follower is an ideal candidate for \textbf{succession} (see \ref{ch:legacy-engine}). When your character retires, falls, or ascends, this contact can inherit your faction standing, one unresolved burden, and a Legacy Bond. Choose a contact you would trust to carry your story forward.
\end{tcolorbox}

\section{Common Background Categories}
\index{backgrounds!categories}

\subsection*{Social Class Backgrounds}
\index{backgrounds!social}\index{aristocratic}\index{merchant}\index{common folk}

\paragraph{Aristocratic House Scion}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Court-Announced}, \textit{Heraldic Right}
\item \textbf{Skills}: Sway, Command, Lore (customs)
\item \textbf{Signature Contact}: Family steward (Cap~1)
\item \textbf{Background Boon}: Once/session in a formal setting, treat Social Position as one step better for an exchange.
\item \textbf{Obligation Seeds}: Marriage alliances, estate audits, house feuds
\item \textbf{Challenges}: Intrigue clocks, reputation stakes
\end{itemize}

\paragraph{Merchant Factor}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Guild-Ledgered}, \textit{Caravan Rights}
\item \textbf{Skills}: Sway (negotiation), Investigation (accounts), Lore (markets)
\item \textbf{Signature Contact}: Bookkeeper or quartermaster (Cap~1)
\item \textbf{Background Boon}: Once/session when buying/selling or logistics, reduce DV by 1 or gain +1d.
\item \textbf{Obligation Seeds}: Notes due, partner demands, embargoes
\end{itemize}

\paragraph{Common Folk}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Neighborhood Trust}, \textit{Guild-Friend}
\item \textbf{Skills}: Craft (one), Endurance, Streetwise or Nature
\item \textbf{Signature Contact}: Elder, foreman, or midwife (Cap~1)
\item \textbf{Background Boon}: Once/session among familiar folk, convert a failure on routine labor to a success-with-cost.
\item \textbf{Obligation Seeds}: Communal favors, festival work, mutual aid
\end{itemize}

\subsection*{Professional Backgrounds}
\index{backgrounds!professional}\index{military}\index{academic}\index{criminal}

\paragraph{Marcher Veteran}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Veteran-of-the-Marches}, \textit{Muster Papers}
\item \textbf{Skills}: Melee or Ranged, Tactics, Endurance
\item \textbf{Signature Contact}: Old sergeant (Cap~1)
\item \textbf{Background Boon}: Once/session in organized violence, ignore the first point of situational Disadvantage from confusion or fear.
\item \textbf{Obligation Seeds}: Recall to colors, feud legacies, back-pay claims
\end{itemize}

\paragraph{Academic of the Dialogues}
\begin{itemize}
\item \textbf{Access Tags}: \textit{School-Sealed}, \textit{Reading Privileges}
\item \textbf{Skills}: Lore, Investigation, Sway (debate)
\item \textbf{Signature Contact}: Lecturer or archivist (Cap~1)
\item \textbf{Background Boon}: Once/session in research or disputation, ask one clarifying question the GM must answer plainly or point to the shelf that does.
\item \textbf{Obligation Seeds}: Citations owed, controversial thesis, patron politics
\end{itemize}

\paragraph{Underworld Hand}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Harbor Handsigns}, \textit{Black Banner Cant}
\item \textbf{Skills}: Stealth, Deception, Streetwise
\item \textbf{Signature Contact}: Fence or fixer (Cap~1)
\item \textbf{Background Boon}: Once/session when using a \emph{cant} (secret coded speech), gain +1d to set up or slip out.
\item \textbf{Obligation Seeds}: Unpaid favors, heat with a rival crew, burned safehouse
\end{itemize}

\subsection*{Regional Backgrounds}
\index{backgrounds!regional}\index{urban}\index{rural}\index{frontier}

\paragraph{Urban Child of Silkstrand}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Street-Stall Rights}, \textit{Night-Patrol Familiar}
\item \textbf{Skills}: Streetwise, Perception, Sway (fast talk)
\item \textbf{Signature Contact}: Watch-lantern or stall-keeper (Cap~1)
\item \textbf{Background Boon}: Once/session in a dense city, you know a \emph{shortcut or rumor} that reduces time/cost by one step.
\end{itemize}

\paragraph{Rural Valewood Forager}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Green Courts Token}, \textit{River-Ferry Nod}
\item \textbf{Skills}: Nature, Survival, Lore (spirits)
\item \textbf{Signature Contact}: Leaf-warden (Cap~1)
\item \textbf{Background Boon}: Once/session in the wild, downgrade a Travel complication by one step.
\end{itemize}

\paragraph{Frontier Marcher Homesteader}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Land-Grant Papers}, \textit{Border Moot Voice}
\item \textbf{Skills}: Craft (build), Endurance, Tactics or Diplomacy
\item \textbf{Signature Contact}: Neighboring elder (Cap~1)
\item \textbf{Background Boon}: Once/session at a contested site, shift Position up by one step when defending what's yours.
\end{itemize}

\section{World-Linked Backgrounds}
\index{backgrounds!cultures}\index{cultures!backgrounds}

\paragraph{Aeler Charter Apprentice}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Charter-Stamped}, \textit{Stone-Speech Initiate}
\item \textbf{Skills}: Craft, Lore (engineering), Investigation (plans)
\item \textbf{Signature Contact}: Oathscribe (Cap~1)
\item \textbf{Boon}: Once/session around structures, ignore the first \emph{environmental} complication from collapse or strain.
\end{itemize}

\paragraph{Mistlands Bell-Keeper}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Bell-Law}, \textit{Causeway Rights}
\item \textbf{Skills}: Perception, Lore (omens), Survival (fens)
\item \textbf{Signature Contact}: Dusk-warden (Cap~1)
\item \textbf{Boon}: Once/session, ring a lawful pattern to reframe a fog scene's Position from \emph{Controlled} to \emph{Dominant} for one exchange.
\end{itemize}

\paragraph{Ykrul Wind-Runner}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Wind-Knot Tokens}, \textit{Banner Hospitality}
\item \textbf{Skills}: Survival (steppe), Ranged or Riding, Sway (customs)
\item \textbf{Signature Contact}: Banner-aunt (Cap~1)
\item \textbf{Boon}: Once/session in open country, coordinate allies at distance; they gain a one-time \emph{Boon} if they can hear your call.
\end{itemize}

\paragraph{Kahfagian Dock-Priest Acolyte}
\begin{itemize}
\item \textbf{Access Tags}: \textit{Harbor Blessings}, \textit{Beacon Rights}
\item \textbf{Skills}: Lore (sea rites), Sway (crews), Investigation (cargo)
\item \textbf{Signature Contact}: Beacon-keeper (Cap~1)
\item \textbf{Boon}: Once/session at sea or harbor, treat one navigation check as if aided by favorable weather (DV--1).
\end{itemize}

\section{Background Benefits and Limits}
\index{backgrounds!benefits}\index{backgrounds!limitations}

\subsection*{Typical Benefits}
\begin{itemize}
\item \textbf{Skill Access}: Strong narrative justification for certain tests
\item \textbf{Resource Access}: Borrowed tools, rooms, shrines, or records (scene-time instead of XP; see Assets, \S\ref{ch:assets-followers})
\item \textbf{Social Access}: Entry to circles your tag recognizes
\item \textbf{Knowledge}: Clues, maps, or customs that lower DV by 1
\item \textbf{Reputation}: Advantage on \emph{first impressions} in-tag
\end{itemize}

\subsection*{Common Limitations}
\begin{itemize}
\item \textbf{Duties}: Responsibilities to house, guild, banner, or court
\item \textbf{Restrictions}: Codes, taboos, or public expectations
\item \textbf{Enemies}: Rival houses, crews, or schools
\item \textbf{Secrets}: Leverage others can pull
\item \textbf{Dependencies}: Loss of standing if you neglect ties
\end{itemize}

\section{Custom Backgrounds}
\index{backgrounds!creating}

\subsection*{Build Procedure (Player \& GM)}
Pick \textbf{1 Access Tag}, \textbf{1 Signature Contact}, \textbf{1 Boon}, and \textbf{1 Obligation Seed}. Name places and people.
\begin{itemize}
\item \textbf{Origin}: Where were you taught/raised/forged?
\item \textbf{Training}: Which \emph{two} skills does this background naturally support?
\item \textbf{Experiences}: One memory that still changes how you act
\item \textbf{Relationships}: One ally, one rival; both want something
\item \textbf{Values}: What won't you let slide?
\end{itemize}

\subsection*{Mechanical Considerations}
\index{backgrounds!mechanics mapping}
\begin{itemize}
\item \textbf{Boon Scope}: DV--1 or +1d once/session when strongly on-theme
\item \textbf{Contact Use}: +1d as assist; cannot take independent actions unless purchased as a \emph{Follower} (Cap$^2$ XP; \S\ref{ch:assets-followers})
\item \textbf{Assets}: Borrowing an institutional asset costs \emph{scene time} (counts as upkeep for that asset this downtime) but \(0\)~XP
\item \textbf{Languages}: Choose one tongue linked to the background (see Chapter \ref{ch:languages} in the full rules); at least \emph{Conversational}
\end{itemize}

\section{Integration with Development}
\index{backgrounds!integration}\index{advancement!backgrounds}

\subsection*{Early Career}
\begin{itemize}
\item Apprentice, recruit, junior clerk, initiate bell-keeper
\item Boon is modest; obligation triggers are frequent but small
\end{itemize}

\subsection*{Mid-Career}
\begin{itemize}
\item Foreman, sergeant, factor, lecturer
\item Add a \textbf{second Access Tag} or elevate contact to Cap~2 by \emph{paying XP} for a follower
\end{itemize}

\subsection*{Late Career}
\begin{itemize}
\item Master, banner-aunt, beacon-warden, arbiter
\item Background Boon may become a \emph{scene-long permission} at GM approval tied to a vow or festival
\end{itemize}

\section{Background Combinations}
\index{backgrounds!combinations}

\paragraph{Sequential}
\begin{itemize}
\item \textbf{Veteran $\rightarrow$ City Watch}: Access shifts from \textit{Muster Papers} to \textit{Badge Rights}
\item \textbf{Scholar $\rightarrow$ Diplomat}: Keep libraries, gain court
\end{itemize}

\paragraph{Parallel}
\begin{itemize}
\item \textbf{Noble Merchant}: Both \textit{Court-Announced} and \textit{Guild-Ledgered}; responsibilities multiply
\end{itemize}

\paragraph{Contrasting}
\begin{itemize}
\item \textbf{Criminal Penitent}: Underworld ties vs.\ temple vows; great hooks, sharp bargains
\end{itemize}

\section{Using Backgrounds in Play}
\index{backgrounds!in play}

\subsection*{Player Applications}
\begin{itemize}
\item Name your places: the specific stall, shrine, or yard you know
\item Cash your Boon when you leverage that familiarity
\item Call on your contact to assist or to introduce a new lead
\end{itemize}

\subsection*{GM Applications}
\begin{itemize}
\item Offer choices where background allegiance matters
\item Turn Access Tags into \emph{gates} and \emph{keys}
\item Spend Story Beats (\S\ref{ch:world-interaction}) to \emph{complicate the old ties}
\end{itemize}

\section{Obligations \& Evolution}
\label{sec:bg-obligation}
\index{backgrounds!obligations}\index{backgrounds!evolution}

\subsection*{Obligation Clock (4)}
When filled by misses, bargains, or story, something is \emph{due}.
\begin{description}
\item[1--2] Minor favor, time, or tithe
\item[3] Significant service, costly honesty, dangerous escort
\item[4] Defining ordeal: duel, audit, rite, or trial
\end{description}
Clear segments by honoring duties on-screen, sacrificing resources, or changing affiliation (which creates new clocks).

\subsection*{Enhancement \& Transformation}
\begin{itemize}
\item \textbf{Enhance}: Add a tag or upgrade the contact via XP (Follower rules)
\item \textbf{Transform}: Rewrite the background after a major arc; keep one tag, swap the rest, reset the clock
\item \textbf{Resolution}: Retire a background after paying its \emph{price}; gain a \emph{Legacy Note}—a small permanent permission
\end{itemize}

\section{Tables \& Tools}
\index{backgrounds!tables}\index{random tables}

\subsection*{Quick Background Seeds}
\small
\begin{longtable}{ll}
\toprule
d10 & Seed \\
\midrule
1 & Failed heir with an honest steward and a dishonest cousin \\
2 & Caravan accountant who knows three border tariffs by heart \\
3 & Bell-tender whose dusk pattern saved a village once \\
4 & Charter apprentice who broke a measure and hid it \\
5 & Ykrul courier who owes horses to two banners \\
6 & Dock acolyte who lit the beacon in a storm against orders \\
\bottomrule
\end{longtable}
\normalsize

\subsection*{GM SB Menu: Background Complications}
\index{complications!background}
\begin{itemize}
\item \textbf{1 SB}: A rival claims your tag is invalid here; prove it or pay time
\item \textbf{2 SB}: Your contact is \emph{busy} or \emph{watched}; help will cost you exposure
\item \textbf{3 SB}: Obligation advances; an audit, rite, or muster arrives early
\item \textbf{4 SB}: An enemy weaponizes your background—publicly
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Background Quick Reference,fonttitle=\bfseries]
\textbf{At Creation}
\begin{itemize}
\item Choose: 1 \emph{Access Tag}, 1 \emph{Signature Contact} (Cap~1 assist), 1 \emph{Boon} (once/session), 1 \emph{Obligation Seed}.
\item Pick 1 language tied to the background at \emph{Conversational} or better (Chapter \ref{ch:languages}).
\end{itemize}

\textbf{In Play}
\begin{itemize}
\item Invoke tags to open doors or shift Position.
\item Call your contact for +1d assist (no independent actions).
\item Track the \emph{Obligation Clock} (4). Pay it on-screen.
\end{itemize}

\textbf{Advancing}
\begin{itemize}
\item Upgrade contact to a \emph{Follower} by paying XP (Cap$^2$).
\item Add a second tag after a major arc or training.
\end{itemize}
\end{tcolorbox}

\section{Worked Examples}
\index{backgrounds!examples}

\subsection*{House Shield (Aristocratic-Military Hybrid)}
\begin{itemize}
\item \textbf{Tags}: \textit{Court-Announced}, \textit{Veteran-of-the-Marches}
\item \textbf{Skills}: Command, Melee, Lore (protocol)
\item \textbf{Contact}: House steward (Cap~1)
\item \textbf{Boon}: Once/session while guarding a principal, treat one consequence as one step less severe.
\item \textbf{Hook}: Duel clock from an old slight; an audit of campaign spoils
\end{itemize}

\subsection*{Ledger Lantern (Merchant-Temple Bridge)}
\begin{itemize}
\item \textbf{Tags}: \textit{Guild-Ledgered}, \textit{Beacon Rights}
\item \textbf{Skills}: Sway, Investigation (accounts), Lore (sea rites)
\item \textbf{Contact}: Beacon-keeper (Cap~1)
\item \textbf{Boon}: Once/session, secure safe harbor or fair rate; DV--1.
\item \textbf{Hook}: Rivals whisper of skimming tithe; prove the books
\end{itemize}

\subsection*{Fog Warden (Mistlands Civic) }
\begin{itemize}
\item \textbf{Tags}: \textit{Bell-Law}, \textit{Causeway Rights}
\item \textbf{Skills}: Perception, Survival, Sway (local law)
\item \textbf{Contact}: Dusk-warden (Cap~1)
\item \textbf{Boon}: Once/session in mist, choose \emph{route-or-safety}: negate a travel delay or downgrade a hazard.
\item \textbf{Hook}: Someone silences the bells; obligation advances
\end{itemize}

\bigskip
\noindent\emph{Let your background do work for you. Speak its language, call its favors, and pay its dues in play.} The world will answer.

\section{Player Bonds: Ties That Bind}
\index{bonds}\index{player relationships}

Your background shapes your past, but the bonds you forge with other player characters shape your future. Bonds are the shared history, favors owed, rivalries, or affections that connect you to one another. They are not merely narrative flavor—they are mechanical resources that allow you to shift Boons, aid each other in desperate moments, and unlock powerful teamwork options late in the campaign.

\subsection{What Bonds Represent}
A bond is a **player-to-player agreement** about how your characters relate. It must be:
\begin{itemize}
\item \textbf{Mutual}: Both players agree on its nature.
\item \textbf{Specific}: It names a concrete event or recurring dynamic.
\item \textbf{Consequential}: It should create roleplaying opportunities and occasional friction.
\end{itemize}

\paragraph{Examples of Bonds}
\begin{enumerate}
\item \textbf{Oath of Shared Vengeance}: “We both lost someone to the same enemy. We swore to see them fall together.”
\item \textbf{Favor of Blood}: “You saved my life when I was helpless. I owe you a favor I cannot erase.”
\item \textbf{Secret Shared}: “We know a truth that could destroy a powerful person. Neither of us can ever speak it alone.”
\item \textbf{Rivalry of Equals}: “We constantly compete, but when outsiders threaten, we fight back-to-back without hesitation.”
\item \textbf{Foster–Kin}: “Our families bound us by rite. We are not siblings by blood, but by custom.”
\item \textbf{Old Campaigners}: “We marched together in the same condotta. We’ve seen each other bleed and lie.”
\end{enumerate}

\subsection{Mechanics of Bonds}

\subsubsection*{Gaining Boons Through Bonds}
Once per bond **per session**, when you take a significant action to aid the other character and explicitly reference your bond in an **intricate description** (see Description Ladder, §\ref{ch:core-mechanics}), you both gain \textbf{1 Boon} after the action resolves.

\textit{You cannot gain this Boon for basic assistance rolls or minor favors. The aid must involve meaningful risk or sacrifice.}

\subsubsection*{Re‑allocating Boons (Late Game)}
At **Tier III or higher** (90+ XP), bonds become a channel for resource sharing. Once per scene, you may transfer **up to 2 Boons** from your pool to a bonded ally as a free action. The transfer must be narrated (“I toss them my last flare,” “I steady their aim with a whispered cue”). This does not count as your action for the turn.

This represents deep trust and practiced coordination—the kind only earned through shared ordeals.

\subsubsection*{Bonded Assistance}
When you assist a bonded ally on a roll, the first Boon you spend to assist (see Assistance, §\ref{ch:core-mechanics}) is **refunded** if the roll succeeds. (You still mark any Fatigue or other costs normally.)

\subsection{Creating Bonds at Character Creation}

At campaign start, each player should establish **two bonds** with other player characters. Use the tables below for inspiration, or invent your own.

\paragraph{Bond Triggers (roll d8 or choose)}
\begin{enumerate}
\item \textbf{Last Stand}: “We held a breach together when no one else would.”
\item \textbf{Betrayal Forgiven}: “You could have left me, but you didn’t.”
\item \textbf{Stolen Future}: “We both lost everything in the same fire.”
\item \textbf{Shared Geas}: “A power bound us to the same task. We cannot complete it alone.”
\item \textbf{Inherited Feud}: “Our families hate each other, but we refuse to carry that hate.”
\item \textbf{Whisper Witness}: “We overheard a secret that would end us if it got out.”
\item \textbf{Trial by Ordeal}: “We survived a rite that killed the others.”
\item \textbf{Empty Promise}: “We made a vow we half‑meant, but now it has teeth.”
\end{enumerate}

\paragraph{Bond Flavors (roll d8 or choose)}
\begin{enumerate}
\item \textbf{Protective}: “I will not let them fall while I draw breath.”
\item \textbf{Resentful but Reliable}: “I don’t like them, but I owe them.”
\item \textbf{Playful Rivalry}: “They beat me at cards every night. I’ll get them back when it counts.”
\item \textbf{Devotion}: “They are the reason I still believe in anything.”
\item \textbf{Curiosity}: “I don’t understand them, but I need to.”
\item \textbf{Dependence}: “Without them, I would have drowned in my own mistakes.”
\item \textbf{Inspiration}: “They make me want to be better.”
\item \textbf{Doom–Bound}: “The omens say one of us will betray the other. I will prove them wrong.”
\end{enumerate}

\subsection{Evolving Bonds}

Bonds are not static. At the end of a major story arc (GM discretion), you may:
\begin{itemize}
\item \textbf{Deepen a bond}: Add a second specific trigger. That bond now grants +1 Boon instead of 1 when invoked (still once per session).
\item \textbf{Sever a bond}: Replace it with a new bond with the same player, or with a different player if the narrative supports it. Severing a bond before its natural conclusion generates \textbf{2 SB} for the GM (the rupture has consequences).
\item \textbf{Transform a bond}: Change its flavor (e.g., from “Resentful” to “Respectful”) and reset its mechanical benefits as if new.
\end{itemize}

\subsection{Bonds and the Late Game}

As characters gain influence (Tier IV–V), bonds become the primary currency of high‑stakes teamwork. Boon transfers allow one character to fuel another’s desperate rite or finishing blow. GMs are encouraged to present challenges where no single character can succeed alone—where only bonded pairs or trios can split responsibilities, cover weaknesses, and share the load.

\paragraph{Example:} The party’s Cantor has 4 Boons but is cornered and low on Fatigue. The Fighter (bonded to the Cantor) transfers 2 Boons as a free action, then uses their own action to defend. The Cantor spends those Boons to reroll a failed casting, saving the party from a collapsing ceiling.

\subsection{Bond Hooks for the Tables Below}

When you choose a background from the previous sections, consider how your bond with another PC might connect to your signature contact, obligation seed, or access tag.

\begin{itemize}
\item \textbf{Aristocratic House Scion + Marcher Veteran}: “Your family hired my company. I still blame you for the lost patrol.”
\item \textbf{Underworld Hand + Urban Child}: “We grew up in the same tenement. You kept my secrets; I kept yours.”
\item \textbf{Aeler Charter Apprentice + Mistlands Bell-Keeper}: “We met in the high passes. Your bell saved me from a fog wraith; my stone‑lore saved you from a collapsed gallery.”
\end{itemize}

\section{Bonds Quick Reference}
\begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,title=Bonds Summary,fonttitle=\bfseries]
\textbf{At Creation:}
\begin{itemize}
\item Establish 2 bonds with other PCs.
\item Define trigger / flavor / mutual benefit.
\end{itemize}

\textbf{Per Session:}
\begin{itemize}
\item Once per bond per session, invoke with intricate description → both gain 1 Boon.
\end{itemize}

\textbf{Tier III+ (90+ XP):}
\begin{itemize}
\item Transfer up to 2 Boons to bonded ally as free action (once per scene).
\end{itemize}

\textbf{Bonded Assistance:}
\begin{itemize}
\item First Boon spent to assist a bonded ally is refunded on a successful roll.
\end{itemize}

\textbf{Evolving:}
\begin{itemize}
\item Deepen, sever, or transform after major arcs.
\end{itemize}
\end{tcolorbox}