% ============================================================
% Chapter 9: Your First Magical Character
% ============================================================
\chapter{Your First Magical Character}
\label{ch:first-magical}
\index{magic!first character}

So you want to play a magic-user. Good choice---but \textsc{Fate's Edge} offers \textbf{six different magical paths}. Jumping into all of them at once is a recipe for confusion. This guide helps you pick a path and then walks you through building a \textbf{single, focused magical character} that is powerful, fun, and manageable.

\section*{Quick Reference: Magic Paths at a Glance}
\index{magic!paths comparison}

Use this table to choose a path that fits your preferred complexity and play style. \textbf{New players: start with Runekeeper.}

\begin{center}
\small
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{lllll}
\toprule
\textbf{Path} & \textbf{Complexity} & \textbf{Cost Type} & \textbf{Risk} & \textbf{Best For} \\
\midrule
Runekeeper & Low & Obligation & Patron demands & First-timers, reliable magic \\
Hedge Gifts & Low & None (Talent) & None & Anyone wanting a magical touch \\
Familiar Only & Low & Obligation (Patron's Gift) & Patron demands & Companionship + utility \\
Cantor & Medium & Corruption & Social fallout, madness & Bards, performers, risk-takers \\
Invoker & Medium--High & Obligation + ritual time & Slow casting, Symbol breaks & Ritual specialists, planners \\
Summoner & High & Leash + Boons & Spirit breaks free & Pet-class fans, tactical players \\
Free Caster & High & Backlash (Fatigue/Harm) & Unpredictable effects & Improv players, wild magic fans \\
Psionics (optional) & Medium & Mental Strain & Mental strain, social distrust & Mind-power enthusiasts \\
\bottomrule
\end{tabular}
\end{center}

\noindent
\textbf{All paths are viable.} The table only shows the \textit{relative} complexity and risk.

If you are still unsure, follow the recommendation below---you can always learn a second path later (see \ref{ch:magic}).

\section{Step 0: The Recommendation}
\index{Runekeeper!as first character}

If this is your first magical character, play a \textbf{Runekeeper}. Why?

\begin{itemize}
  \item \textbf{One Patron:} You swear to a single Patron. No juggling multiple Symbols or conflicting obligations.
  \item \textbf{Clear costs:} Every Rite costs Obligation. Track one number.
  \item \textbf{Support from your Thiasos:} Your familiar is a built-in roleplay hook and mechanical aid.
\end{itemize}

You can explore Invokers, Cantors, Witches, Psions, or Summoners later. Start simple.

\subsection*{Which Magic Path Fits You?}
\index{magic!player fit}

Not every player enjoys the same kind of magic. Use this table to find the path that matches how you like to play.

\begin{center}
\small
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l X X}
\toprule
\textbf{If you want\dots} & \textbf{Choose this path} & \textbf{Why it fits} \\
\midrule
Reliable, straightforward magic with clear costs & \textbf{Runekeeper} & One Patron, one resource (Obligation). No surprises. \\
A magical touch without committing to a full path & \textbf{Hedge Gifts} & Spend 4 XP for a few no-roll tricks. Retrain later. \\
A companion with a little magical utility & \textbf{Familiar Only} & Get Patron's Gift and a follower for 2 XP. Costs Obligation per use. \\
To sing, perform, and risk losing yourself & \textbf{Cantor} & Corruption timer creates drama; great for bards. \\
Rituals, preparation, and borrowing power & \textbf{Invoker} & Slow but flexible; collect Symbols like tools. \\
A summoned creature that fights for you & \textbf{Summoner} & Tactical pet class; manage the Leash timer. \\
Improvised magic on the fly & \textbf{Free Caster} & TAGS system lets you invent spells; high risk. \\
Mind powers and subtle influence (optional) & \textbf{Psionics} & No Patron, but mental strain and social suspicion. \\
\bottomrule
\end{tabularx}
\end{center}

\noindent
Still unsure? \textbf{Let's start with Runekeeper.} You can always learn a second path later (see \ref{ch:magic}).

\section{Two Low-Cost Alternatives}

Before committing to a full magical path, consider these two low-cost options. Both give you a magical feel without the complexity of Rites, Corruption, or a Leash.

\subsection{Hedge Gifts: Magic Without a Path}
\index{Hedge Gifts}
\index{magic!low-cost alternatives}

\textbf{Cost:} 4 XP (Craft of the Hedge talent)

Hedge Gifts are simple, no-roll magical abilities. A character with the \textbf{Craft of the Hedge} talent learns two Hedge Gifts of their choice. These are small, reliable effects that never require dice rolls and never incur Obligation.

\textbf{Why take Hedge Gifts?}
\begin{itemize}
  \item \textbf{Cheap:} 4 XP gives you two useful abilities.
  \item \textbf{No resource tracking:} No Obligation, Corruption, or Backlash.
  \item \textbf{Retrain later:} If you want to commit to a full magical path, you can retrain this talent during downtime.
  \item \textbf{Perfect for non-magical characters:} A warrior with a few Hedge Gifts is still a warrior---just one who knows a few tricks.
\end{itemize}

\textbf{Sample Hedge Gifts:}
\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Gift} & \textbf{Effect} \\
\midrule
Steady Hand & Once per scene, touch someone and remove 1 Fatigue. \\
Knot of Favor & Tie a knot; allies gain +1 die to one type of roll for the scene. \\
Warm Hand & Touch someone; their next physical action gains +1 die or suffers -1 die. \\
Hearth-Sense & Ask the GM one yes/no question about a threshold or promise. \\
Salt Line & Pour a salt line; spirits must test to cross. \\
The Unlit Candle & Extinguish a light source you can see. \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Build example (32 XP):}
\begin{itemize}
  \item Attributes: Spirit 3 (15 XP), Wits 2 (6 XP) $\rightarrow$ 21 XP
  \item Skills: Lore 2 (6 XP), Insight 1 (2 XP) $\rightarrow$ 8 XP
  \item Talent: Craft of the Hedge (4 XP) $\rightarrow$ 4 XP
  \item \textbf{Total:} 21 + 8 + 4 = \textbf{33 XP}
\end{itemize}

\subsection{Familiar Only: Thiasos Without a Codex}
\index{Familiar}
\index{Thiasos}
\index{Patron's Gift}
\index{Obligation!Patron's Gift}

\textbf{Cost:} 2 XP (Familiar talent)

You can take the Familiar talent (Thiasos) \textbf{without} taking a Codex. This gives you:

\begin{itemize}
  \item A small creature bound to you (raven, candle flame, fungal creature, etc.)
  \item Access to \textbf{Patron's Gift}: Once per scene, imbue a held item with +1 Melee and +1 Thematic skill for the scene.
  \item A narrative companion and roleplay hook.
  \item \textbf{No Rite Obligation} (you don't know any Rites).
\end{itemize}

\begin{tcolorbox}[colback=yellow!5!white,colframe=yellow!75!black,title=Important: Patron's Gift Costs Obligation,breakable]
\index{Obligation!Patron's Gift}
\textbf{Patron's Gift is not free.} Each time you invoke it, you mark \textbf{+1 Obligation} to your Patron.

Your Obligation Capacity is \textbf{Spirit + Presence}. If you exceed capacity, you mark Fatigue (1 per segment over). You can clear Obligation during downtime by performing an Act of Service for your Patron.

This means:
\begin{itemize}
  \item A Spirit 3, Presence 1 character can invoke Patron's Gift 4 times before hitting Fatigue.
  \item You're still managing a resource---it's just simpler than tracking Rites, Corruption, or a Leash.
  \item You can buy a Codex later to become a full Runekeeper.
\end{itemize}
\end{tcolorbox}

\textbf{Why take Familiar only?}
\begin{itemize}
  \item \textbf{Cheapest magical entry:} 2 XP for a companion and a useful scene-long buff.
  \item \textbf{Minimal resource tracking:} Only Obligation from Patron's Gift (no Rites, no Corruption).
  \item \textbf{Retrain later:} If you want to become a full Runekeeper, buy a Codex (4 XP) during downtime.
  \item \textbf{Follower potential:} Your Thiasos can act as a Cap 1 follower, scouting, carrying small items, or delivering messages.
\end{itemize}

\textbf{Build example (32 XP):}
\begin{itemize}
  \item Attributes: Spirit 3 (15 XP), Wits 2 (6 XP) $\rightarrow$ 21 XP
  \item Skills: Lore 2 (6 XP), Survival 1 (2 XP), Insight 1 (2 XP) $\rightarrow$ 10 XP
  \item Talent: Familiar (2 XP) $\rightarrow$ 2 XP
  \item \textbf{Total:} 21 + 10 + 2 = \textbf{33 XP}
\end{itemize}

\textbf{Patron's Gift in action:}
Your familiar is a brass compass that hums when you're on the right path (Traveler). Once per scene, you touch a map or a walking stick and gain +1 Melee and +1 Navigation for the scene. You mark +1 Obligation. No Rites. No Corruption. Just a small edge---and a small debt.

\subsection{Comparison: Hedge Gifts vs. Familiar Only}

\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Feature} & \textbf{Hedge Gifts} & \textbf{Familiar Only} \\
\midrule
Cost & 4 XP & 2 XP \\
Resource Tracked & None & Obligation (Patron's Gift) \\
Roll Required & No & No (Patron's Gift is an action) \\
Can Retrain & Yes & Yes (buy Codex later) \\
Companion & No & Yes (Thiasos) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Retraining Later}
\index{retraining}
\index{downtime!retraining}

Both options are \textbf{low-cost investments} that can be retrained later.

\begin{itemize}
  \item \textbf{Hedge Gifts:} During downtime, you can retrain Craft of the Hedge into another talent of equal or lesser cost (4 XP or less). The Hedge Gifts are lost, but you can buy a Codex or Patron's Symbol instead.
  \item \textbf{Familiar Only:} During downtime, you can buy a Codex (4 XP) and become a full Runekeeper. Your Thiasos remains; you now have access to Rites and still track Obligation.
\end{itemize}

This means you can start with a magical touch and grow into a full caster as you understand the system better.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=The Gray Wanderer's Advice,breakable]
``You don't need to be a wizard to have a little magic. A handful of Hedge Gifts or a familiar who nudges fate is enough to make a character feel touched by the supernatural. But remember: even a small favor from a Patron is a debt. Mark your Obligation. Pay it off in stories and service. The road will teach you the rest.''
\end{tcolorbox}

\section{Step 1: Choose a Patron (For Runekeepers and Familiar Only)}
\index{Patrons!choosing a first Patron}

Your Patron defines your Rites, your Obligation, and your role in the story. Pick one that matches your character concept.

\begin{center}
\small
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lll}
\toprule
\textbf{Patron} & \textbf{Domain} & \textbf{Character Concept} \\
\midrule
The Traveler & Roads, journeys, safe passage & Guide, courier, scout \\
Oath of Flame \& Light & Dawn, vows, protection & Paladin, healer, knight \\
Grimmir & Nature, seasons & Druid, ranger, hedge-witch \\
Ikasha & Shadow, secrets, thresholds & Spy, infiltrator, night-walker \\
The Witness & Truth, memory, revelation & Investigator, judge, archivist \\
Mykkiel & Law, covenant, judgment & Templar, inquisitor, notary \\
The Carrion-King & Decay, renewal, compost & Necromancer (gentle), healer \\
\bottomrule
\end{tabular}
\end{center}

\textbf{New player favorite:} \emph{The Traveler}. Low Obligation costs, useful Rites, and a forgiving Patron.

\section{Step 2: Buy Your Core Talents (Runekeeper)}
\index{talents!Runekeeper requirements}

At character creation (30--36XP), a full Runekeeper needs two talents:

\begin{itemize}
  \item \textbf{Familiar (Thiasos) -- 2 XP.} Your bond to your Patron. It might be a raven (Traveler), a candle flame (Oath of Flame), or a small fungal creature (Carrion-King).
  \item \textbf{Codex -- 4 XP.} Your book of Rites. It could be a leather journal, a set of carved slate tablets, or a roll of bark.
\end{itemize}

\textbf{Total so far: 6 XP.} You have 24--28 XP remaining for Attributes, Skills, and optional talents.

\section{Step 3: Set Your Attributes}
\index{Attributes!for Runekeepers}

Runekeepers rely on \textbf{Spirit} (Obligation capacity, Rite effectiveness) and \textbf{Wits} (Lore, Investigation). A simple spread:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Attribute} & \textbf{Rating} & \textbf{Cost (new rating \(\times\)3)} \\
\midrule
Spirit & 3 & 9 XP (1$\to$2 = 6, 2$\to$3 = 9) \\
Wits & 2 & 6 XP (1$\to$2 = 6) \\
Body & 1 & 0 (base) \\
Presence & 1 & 0 (base) \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Total: 15 XP.} You have 9--13 XP remaining.

\section{Step 4: Buy Your Skills}
\index{Skills!for Runekeepers}

You need \textbf{Lore} (to understand your Rites) and at least one social or utility skill.

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Skill} & \textbf{Rating} & \textbf{Cost (new level \(\times\)2)} \\
\midrule
Lore & 2 & 6 XP (0$\to$1 = 2, 1$\to$2 = 4) \\
Insight & 1 & 2 XP (0$\to$1 = 2) \\
Sway & 1 & 2 XP (0$\to$1 = 2) \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Total: 10 XP.} You now have 0--3 XP left---perfect for one minor talent.

\section{Step 5: Choose Your First Rites}
\index{Rites!starting Rites}

You begin with \textbf{two Low Rites} from your Patron. Pick ones that are useful and clear.

\subsection*{Example: The Traveler}
\begin{itemize}
  \item \textbf{Road-Sense (Low, 4 XP):} Gain +1 die to avoid ambushes or delays on a journey. \textit{Use this every travel leg.}
  \item \textbf{Traveler's Boon (Low, 5 XP):} Ignore one level of difficult terrain or bureaucracy for the scene. \textit{Great for crossing rivers or skipping paperwork.}
\end{itemize}

\subsection*{Example: Oath of Flame \& Light}
\begin{itemize}
  \item \textbf{Kindle Vow (Low, 4 XP):} Declare a short vow (e.g., ``I will protect the villagers''). Gain +1 die to actions fulfilling it.
  \item \textbf{Lay on Hands (Low, 5 XP):} Cleanse an affliction or downgrade Harm by 1. \textit{The classic healing Rite.}
\end{itemize}

\section{Step 6: Understand Your Resources}
\index{Obligation!tracking}
\index{Fatigue!tracking}

You track three numbers:

\begin{enumerate}
  \item \textbf{Obligation (to your Patron).} Capacity = Spirit + Presence. For Spirit 3, Presence 1, capacity is 4. Each Rite marks +1 Obligation. When you exceed capacity, mark 1 Fatigue per extra segment.
  \item \textbf{Fatigue.} Each point worsens Position. When Fatigue fills your Body track (Body 1 = 1 segment), convert to Harm.
  \item \textbf{Harm.} Harm 1 = --1 die, Harm 2 = --2 dice, Harm 3 = incapacitated.
\end{enumerate}

\textbf{Player's job:} Mark these yourself. Announce when Obligation fills. The GM does not track them for you.

\section{Step 7: Play Your First Session}
\label{sec:first-session-magic}

\subsection*{Before the session}
\begin{itemize}
  \item Write your two Rites on an index card.
  \item Set your Obligation to 0.
  \item Set your Fatigue to 0.
\end{itemize}

\subsection*{During the session}
\begin{enumerate}
  \item \textbf{Invoke a Rite:} Say ``I invoke \emph{Road-Sense}.'' Mark +1 Obligation. Roll Lore + Spirit (or as specified). Apply the effect.
  \item \textbf{Take a Short Rest:} Clear 2 Fatigue.
  \item \textbf{Accept a Patron's demand:} If the GM says ``your Patron requires a task,'' treat it as a story hook, not a punishment.
\end{enumerate}

\subsection*{After the session}
\begin{itemize}
  \item Clear Obligation by performing an \textbf{Act of Service} for your Patron (e.g., The Traveler: guide a lost traveler; Oath of Flame: speak a public vow). Describe the act. The GM will tell you how much Obligation clears.
  \item Spend XP to raise Attributes, Skills, or buy new Rites.
\end{itemize}

\section{Optional: Add a Minor Talent}
\index{talents!minor}

If you have 2--4 XP left, take one of these:

\begin{itemize}
  \item \textbf{Efficient Invocation (4 XP):} Reduce Obligation cost of one specific Low Rite by 1 (minimum 0).
  \item \textbf{Keen Senses (2 XP):} +1 die on perception checks.
  \item \textbf{Silver Tongue (2 XP):} +1 die on persuasion attempts.
\end{itemize}

\section{What to Avoid at First}
\label{sec:avoid-first}

\begin{itemize}
  \item \textbf{Multiple Patrons.} Stick to one. Cross-Resonance (see \ref{ch:magic}) adds complexity you don't need yet.
  \item \textbf{Crack the Seal (Invoker ability).} You are a Runekeeper. Let Invokers handle emergency rituals.
  \item \textbf{Cantor's Corruption.} A great path, but the Corruption timer requires active management. Save it for your second character.
  \item \textbf{Summoning.} The Leash timer and spirit whims table are fun, but they add significant complexity. Start with Runekeeper, then explore Summoning later.
\end{itemize}

\section{Sample Build: Kestra the Traveler's Runekeeper}
\label{sec:sample-build}

\begin{center}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{ll}
\toprule
\textbf{Element} & \textbf{Choice} \\
\midrule
Name & Kestra \\
Patron & The Traveler \\
Attributes & Spirit 3 (9 XP), Wits 2 (6 XP), Body 1, Presence 1 \\
Skills & Lore 2 (6 XP), Insight 1 (2 XP), Sway 1 (2 XP) \\
Talents & Familiar (2 XP), Codex (4 XP) \\
Rites & Road-Sense (4 XP), Traveler's Boon (5 XP) \\
Starting Boons & 3 \\
Obligation Capacity & 4 (Spirit 3 + Presence 1) \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Total XP spent:} 9+6+2+2+2+4+4+5 = 36XP (using the +4 XP from Bonds/Complications).

\medskip
\noindent Kestra is a former caravan guide who swore to the Traveler after being betrayed by her crew. She carries a brass compass (her Thiasos) and a leather journal of route-maps (her Codex). She invokes \emph{Road-Sense} before every journey and \emph{Traveler's Boon} to cross difficult terrain. She tracks her Obligation carefully and pays it off by guiding lost travelers through the mountains.

\section{Sample Build: The Hedge-Witch (Low-Cost Alternative)}
\label{sec:sample-hedge-witch}

\begin{center}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{ll}
\toprule
\textbf{Element} & \textbf{Choice} \\
\midrule
Name & Sariel \\
Patron & None (Hedge Gifts) \\
Attributes & Spirit 3 (9 XP), Wits 2 (6 XP), Body 2 (6 XP) \\
Skills & Lore 2 (6 XP), Survival 1 (2 XP), Insight 1 (2 XP) \\
Talents & Craft of the Hedge (4 XP) \\
Hedge Gifts & Steady Hand, Knot of Favor \\
Starting Boons & 3 \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Total XP spent:} 9+6+6+6+2+2+4 = 35XP (with one Bond or Complication).

\medskip
\noindent Sariel is a village healer who learned a few tricks from her grandmother. She doesn't swear to any Patron---she just knows how to tie knots that hold luck and press hands that soothe pain. If she later wants more magic, she can retrain Craft of the Hedge into Spellcraft or a Codex.

\section{Sample Build: The Familiar-Bonded Scout (Low-Cost Alternative)}
\label{sec:sample-familiar-only}

\begin{center}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{ll}
\toprule
\textbf{Element} & \textbf{Choice} \\
\midrule
Name & Finn \\
Patron & The Traveler (Familiar Only) \\
Attributes & Wits 3 (9 XP), Body 3 (15 XP), Spirit 1, Presence 1 \\
Skills & Survival 2 (6 XP), Stealth 1 (2 XP), Navigation 1 (2 XP) \\
Talents & Familiar (2 XP) \\
Patron's Gift & +1 Melee and +1 Navigation (once per scene). \textbf{Costs +1 Obligation per use.} \\
Starting Boons & 3 \\
Obligation Capacity & 2 (Spirit 1 + Presence 1) \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Total XP spent:} 9+15+6+2+2+2 = 36XP (with one Bond or Complication).

\medskip
\noindent Finn is a scout who carries a raven (his Thiasos) that helps him find safe routes. He has no Rites and no Codex---just the raven's guidance and the Traveler's subtle favor. Each time he calls on that favor, he marks +1 Obligation. With Spirit 1 and Presence 1, his capacity is 2---he can invoke Patron's Gift twice before he starts taking Fatigue. He pays off his Obligation by guiding lost travelers through the hills.

\section{Final Advice}
\label{sec:final-advice-magic}

\begin{tcolorbox}[
  colback=green!5!white,
  colframe=green!75!black,
  title=The One Rule for New Magical Players,
  fonttitle=\bfseries,
  breakable
]
\emph{Spend your Boons. Hoarding them for XP conversion starves the table of drama. A re-rolled failure, an improved Position, or an activated Asset creates a memorable moment. The XP will come from your adventures.}
\end{tcolorbox}

You are now ready to play a magical character---whether a full Runekeeper, a hedge-witch with a few tricks, or a scout with a familiar. When you feel comfortable, read about the other paths (see \ref{ch:magic}). But for now, start simple. The road is long. The magic will grow.

The road is waiting. Pay your toll.