\chapter{Gear, Magic Items, and Crafting}
\label{ch:gear}

\section{The Philosophy of Gear}
In Fate’s Edge, your character’s possessions are described in the fiction, not tracked on a ledger. You have enough coin, barter, or credit to afford:
\begin{itemize}
  \item Ordinary meals, lodging, and travel rations
  \item Basic tools (rope, lantern, writing kit, lockpicks)
  \item Small bribes (a drink, a gift)
  \item Routine transportation (ferry, caravan passage)
\end{itemize}
If the GM ever asks ``Do you have enough?'' the answer is usually \textbf{yes} – unless scarcity is the point of the scene.

\section{Mundane Equipment}
\label{sec:mundane-gear}

\subsection{Weapons and Armor}
Weapons and armor are covered in \S\ref{sec:weapons-armor} (Core Mechanics). Briefly:
\begin{itemize}
  \item \textbf{Light weapons (4 XP):} fast, concealable.
  \item \textbf{Medium weapons (8 XP):} balanced.
  \item \textbf{Heavy weapons (12 XP):} slow, powerful.
  \item \textbf{Light armor (4 XP):} converts 1 Harm → 1 Fatigue.
  \item \textbf{Medium armor (8 XP):} converts 2 Harm → 1 Fatigue, –1 die physical skills.
  \item \textbf{Heavy armor (12 XP):} converts 3 Harm → 2 Fatigue, –2 dice physical skills, no sprint in rough terrain.
\end{itemize}
You may acquire these with XP at character creation or later.

\subsection{Adventuring Gear}
You are assumed to have the following unless you are in a survival situation:
\begin{itemize}
  \item Backpack, waterskin, 1d6 days of rations, a knife, flint and steel, a small lantern or candle
  \item Tools for any skill you have at level 1+ (e.g., lockpicks for Subterfuge, healer’s kit for Medicine)
  \item A set of clothing appropriate to your culture
\end{itemize}
If you need something unusual (a climber’s kit, a disguise, a rare herb), you may:
\begin{itemize}
  \item Spend 1 Boon to declare you already have it (narrative permission).
  \item Spend 1 XP to acquire it (as a one‑time resource).
  \item Acquire it as a minor asset (4 XP) if you want permanent access.
\end{itemize}

\section{Magic Items}
Magic items are not bought with gold. They are acquired by spending XP, completing quests, or crafting. A magic item is a \textbf{permanent asset} that grants a mechanical benefit. It requires \textbf{upkeep} (see below).

\subsection{Cost Basis: Talent Equivalents}
A magic item’s effect is roughly equivalent to a \textbf{Talent} (see \ref{ch:talents}). Use the Talent XP cost as the base price, then apply a \textbf{discount} because items can be lost, stolen, or broken.

\begin{center}
\begin{tabular}{lcc}
\toprule
\textbf{Effect Power} & \textbf{Talent Cost} & \textbf{Magic Item Cost (XP)} \\
\midrule
Minor (e.g., +1 die to a narrow skill once/scene) & 2–3 XP & 2 XP \\
Major (e.g., +1 die to a broad skill always) & 4–6 XP & 4 XP \\
Prestige (e.g., once/session reroll a failed save) & 7–10 XP & 6 XP \\
Epic (e.g., teleport once/day) & 11+ XP & 8 XP \\
\bottomrule
\end{tabular}
\end{center}
\emph{Discount reason:} A Talent is permanently attached to your character; an item can be taken away. The discount also covers the need for upkeep.

\subsection{Sample Magic Items}

\begin{longtable}{p{3.5cm}p{3.5cm}p{7cm}}
\toprule
\textbf{Name} & \textbf{Cost (XP)} & \textbf{Effect} \\
\midrule
Emberfang Dagger & 2 & When drawn, glows faintly if undead are within Near. Once per scene, gain +1 die to detect hidden spirits. \\
Storm Lantern & 4 & Once per session, as an action, create a gust of wind that extinguishes all mundane flames in Near and gives +1 Position to disengage. \\
Cloak of the Unseen & 4 & +1 die to Stealth in dim light. Once per scene, reroll a failed Stealth check. \\
Vial of the Last Breath & 2 (consumable) & Heal 1 Fatigue when drunk. One use. \\
Pendant of the Witness & 6 & Once per session, when you detect a lie, the liar suffers 1 Fatigue (no save). \\
Raven’s Key & 8 & Once per day, as an action, unlock any mundane lock or open any door that is not magically sealed. \\
\bottomrule
\end{longtable}

\subsection{Attunement and Upkeep}
A character may be \textbf{attuned} to up to 3 magic items at a time (attunement takes a short rest). During each downtime, you must pay \textbf{upkeep} for each magic item you keep attuned:
\begin{itemize}
  \item \textbf{Efficient:} Pay XP equal to the item’s cost divided by 3 (rounded up, minimum 1). You spend minimal effort.
  \item \textbf{Intensive:} Pay 1 XP and describe a downtime scene where you maintain the item (cleaning, recharging, appeasing its spirit).
\end{itemize}
If you neglect upkeep, the item becomes \textbf{Neglected} (no benefit until you pay the upkeep retroactively). If neglected for two consecutive downtimes, it becomes \textbf{Compromised} (requires a quest to restore).

\subsection{Losing Magic Items}
If you lose a magic item (theft, destruction, sacrifice), you lose the XP you spent on it. The GM may allow you to recover it as a story hook. You may also choose to ``retire'' an item (sell it, trade it, give it away) and regain half its XP cost (rounded down) to spend on something else.

\section{Crafting Magic Items}
\label{sec:crafting}

Crafting a permanent magic item is a \textbf{campaign‑level project}. It requires:
\begin{itemize}
  \item A character with relevant skills: \textbf{Arcana 3+} and \textbf{Craft 2+} (or appropriate specialty).
  \item A workspace (e.g., a workshop asset, a consecrated forge).
  \item Rare components (obtained through adventuring – the GM will decide).
  \item A \textbf{Downtime Project Timer } (6–12 segments). Each segment requires a scene or a downtime action.
  \item The XP cost of the item (see table above) – paid when the timer is completed.
\end{itemize}

\subsection{Crafting Procedure}
\label{subsec:crafting-procedure}
\begin{enumerate}
  \item \textbf{Design:} Describe what the item does. The GM sets the Tier (Minor/Major/Prestige/Epic) and XP cost.
  \item \textbf{Quest:} The GM frames a short arc to obtain a rare component (e.g., ``a phoenix feather'', ``the heart of a storm elemental'').
  \item \textbf{Downtime Timer :} Roll \textbf{Wits + Craft} or \textbf{Wits + Arcana} each downtime. Each success advances the timer by 1. Partial successes create complications (lost components, attracting attention). On a failure, the timer may tick backwards.
  \item \textbf{Completion:} When the timer fills, you pay the XP and the item is complete. The GM may add a final, optional cost (e.g., ``it also consumes a memory'').
\end{enumerate}

\subsection{Crafting Consumables}
One‑use items (potions, scrolls, runestones) are much easier. Crafting a consumable:
\begin{itemize}
  \item Takes a single downtime action (no timer).
  \item Costs \textbf{1 XP} per consumable (or 2 XP for a powerful one).
  \item Requires an appropriate skill roll (DV 3) – on a failure, the components are wasted.
  \item No upkeep required.
\end{itemize}
Examples: \emph{Healing potion} (restores 1 Fatigue), \emph{Sight scroll} (grants +1 die to Investigation for one scene), \emph{Thunderstone} (ranged attack, Harm 1 to a small area).

\section{Upgrading Mundane Gear}
You may also spend XP to give a mundane item a permanent magical property. Use the same costs as magic items. The process is faster than crafting from scratch (no rare component quest, just a downtime timer of 4–6 segments).

\section{Summary for Players}
\begin{itemize}
  \item Ordinary gear is free – describe what you carry.
  \item Weapons and armor cost XP (light 4, medium 8, heavy 12).
  \item Magic items cost XP based on Talent equivalents (2–8 XP).
  \item Magic items require upkeep (XP or downtime scene) each downtime.
  \item Crafting a magic item is a long project involving a timer, rare components, and XP.
\end{itemize}

\begin{tcolorbox}[colback=green!5!white,colframe=green!70!black,title=The Grey Wanderer’s Tip]
``A magic sword is not a toy. It is a creditor. It demands maintenance, attention, and sometimes a blood‑price. If you cannot afford the upkeep, do not pick it up. The edge is not worth the interest.''
\end{tcolorbox}

\section{Artifacts: Items That Owe You (And You Owe Them)}

\emph{The coin that never spends is the one you don’t remember taking. The blade that never dulls is the one you never truly own.}

Most magic items are tools – you pay XP, you keep them, you maintain them. \textbf{Artifacts} are different. They are not crafted; they are inherited, discovered, or earned through story. An artifact has a will of its own, however faint. It remembers every hand that held it, every oath spoken over it, every debt left unpaid.

\subsection{What Makes an Artifact?}

An artifact is a magic item with three special properties:

\begin{itemize}
  \item \textbf{Obligation.} When you first claim an artifact (or when you use its deeper powers), you mark Obligation to the artifact itself. This Obligation is tracked like Patron Obligation (Capacity = Spirit + Presence). The artifact may demand tasks, sacrifices, or simply that you carry it to places it wishes to go.
  \item \textbf{Rites.} An artifact grants one or two specific Rites (see \ref{ch:world-powers-obligation}) that you can invoke without a Patron, a Codex, or a Symbol. Using a Rite from an artifact costs the normal Obligation (in addition to any Obligation from attuning to the artifact). The Rites are tied to the artifact’s nature – a sword that remembers a betrayed king might grant \textit{Rite of the Oath Unbroken}; a lantern that once lit a heretic’s pyre might grant \textit{Rite of the Revealing Flame}.
  \item \textbf{Indestructible (mostly).} If an artifact is lost, stolen, or destroyed, it reappears intact at the next dawn – somewhere nearby, often in the hand of the one who owes it the most Obligation. The only way to be rid of an artifact permanently is to fulfill all Obligation to it, then pass it to a willing heir (or bury it in a place of true rest, with GM approval).
\end{itemize}

\subsection{No Upkeep, but No Escape}

Unlike ordinary magic items, artifacts require \textbf{no upkeep} (no XP or downtime scenes). They maintain themselves – or rather, they maintain their hold on you. However, you cannot simply stop attuning to an artifact; it will return at dawn until your Obligation to it is cleared.

\begin{itemize}
  \item \textbf{Attunement:} An artifact counts toward your attunement limit (3 items). Breaking attunement is a ritual (downtime scene) that requires a \textbf{Spirit + Lore} roll (DV 4). On a failure, you mark 1 Fatigue and the artifact reappears the next dawn.
  \item \textbf{Obligation Overflow:} If your Obligation to an artifact exceeds your Capacity, you suffer Fatigue as normal (see \S\ref{sec:obligation-overflow}). Additionally, the artifact may manifest – a cold weight in your pack, a whisper in your ear, a reflection that doesn’t match the room.
\end{itemize}

\subsection{Subtle Power, Not Apocalyptic}

Most artifacts are \textbf{subtle}. They do not reshape battlefields or raise armies. A true artifact might:

\begin{itemize}
  \item Allow you to hear the last words of anyone who died holding it (once per session).
  \item Show you the nearest hidden door or secret passage when you spend 1 Boon.
  \item Grant +1 die to all Sway rolls when bargaining with a specific faction (e.g., the Salt Prince’s court).
\end{itemize}

Apocalyptic artifacts – those that could end kingdoms or unmake reality – are almost unheard of. One or two are whispered about in the oldest ledgers, but no living witness has confirmed their existence. The Grey Wanderer themselves is said to carry a shard of the First Oath, but if they do, they have never used it.

\subsection{Acquiring an Artifact}

You cannot buy an artifact with XP. You find them:

\begin{itemize}
  \item In the hoard of a forgotten dragon.
  \item At the bottom of a sunken temple, still clutched by a skeleton.
  \item As an inheritance from a dying mentor (``Take my mother’s mirror. It will show you what I could not face.'').
  \item As a reward from a Patron who wishes to bind you more tightly.
\end{itemize}

The GM determines the artifact’s Rites and the starting Obligation (usually 1–3 segments). You do not pay XP for the artifact itself, but you may need to complete a quest to unlock its full power.

\subsection{Sample Artifacts}

\begin{longtable}{p{4cm}p{3cm}p{9cm}}
\toprule
\textbf{Name} & \textbf{Obligation} & \textbf{Effect} \\
\midrule
\textit{The Copper Bell of Unspoken Truths} & 2 & Once per session, ring the bell. The next lie told within Near range is heard as a discordant chime; you gain +1 die to expose it. \\
\textit{Ledger of the Debtor Saint} & 3 & You may record a single promise in the ledger. While the promise is unfulfilled, you gain +1 die to all rolls that directly advance it. When the promise is fulfilled, clear 1 Obligation. \\
\textit{The Mourner’s Cloak} & 1 & When you pull the hood up, you become invisible to anyone who is actively grieving (GM’s judgment). The invisibility lasts until you speak or attack. \\
\textit{Hearthstone of the Lost Hearth} & 2 & Once per day, as an action, you may set the stone on a threshold. For the next hour, that threshold cannot be crossed by anyone with hostile intent (Resolve DV 4 to resist). \\
\bottomrule
\end{longtable}

\subsubsection{The Kalrantha Amulet (Tier V Artifact)}
A tear‑shaped shard of green glass in a brass frame. Warm to the touch when someone nearby is about to decide.
\begin{itemize}
    \item \textbf{Fate's Gravity:} Once per session, reroll a single die after seeing the result.
    \item \textbf{Anti‑Magic Aura (Near):} Any spell that alters probability, fate, or causality suffers +2 DV within Near range.
    \item \textbf{Story Beats Become Gravity:} When the wearer rolls 1s, those SB are transferred to the next player character who acts. The GM must spend them on inevitable complications for that character's action.
\end{itemize}
For full corruption mechanics and adventure hooks, see \textit{Dhahara: Kingdom of Brass} and \textit{The Way of Silk}.

\subsection{Player Advice}

Artifacts are \textbf{story hooks}, not power-ups. If you find one, ask yourself:
\begin{itemize}
  \item Why does this thing want me to carry it?
  \item What will it demand when my Obligation fills?
  \item Who else wants it – and what will they pay?
\end{itemize}
Embrace the artifact as a co-star in your character’s drama. The Rites it grants are useful; the strings attached are the real treasure.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!70!black,title=The Grey Wanderer’s Ledger]
``I have carried the Copper Bell for thirty years. It has shown me a hundred lies and saved me from a thousand embarrassments. It has also woken me at midnight, demanded I ring it at a stranger’s door, and once caused a magistrate’s wig to catch fire. I would not trade it for a kingdom – but I would trade it for a single night of silence.''
\end{tcolorbox}