\section{The Witch Hunter}
\index{character concepts!witch hunter}\index{witch hunter}

\textbf{Concept}: A sanctioned tracker of illicit rites and rogue talents. Part detective, part exorcist, part collector of debts owed to the law. Witch hunters serve mortal authorities, temple courts, or---most dangerously---a divine Patron who demands purity.

\textbf{Typical Inspiration}: Inquisitors, occult constables, hedge-wardens, contract witch-finders, zealots of the Inquisitor Prime.

\textbf{Mechanical Foundation}:
\begin{itemize}
  \item \textbf{Attributes}: Wits 4, Spirit 3 (or substitute Body 3 for a more martial hunter)
  \item \textbf{Skills}: Investigation 3, Lore (occult) 2, Melee 2, Insight 1, Command 1
  \item \textbf{Talents}: 
    \begin{itemize}
      \item \emph{Witch-Sign Reader} -- spot residue of magical workings; gain +1 die on Investigation vs. occult scenes.
      \item \emph{Cold Iron Practice} -- gain +1 die when disrupting active magic or attacking a supernatural creature with cold iron.
      \item \emph{Calm Under Oath} -- once per scene, resist fear, panic, or intimidation (reroll a failed Spirit + Resolve test).
    \end{itemize}
  \item \textbf{Assets}: Writ of limited authority (opens doors, closes some), temple archive informant, a cold iron weapon (Minor Asset)
  \item \textbf{Affiliation}: Choose one order -- Chain-Lanterns (Aeler), Spirit-Shield (Mistlands), Sum-Abjurers (Thepyrgos), Break-Reins (Vilikari), Silk/Shed Vigil (Silkstrand/Zakov), Wake-Judges (Ecktoria), or the \textbf{Inquisition of the Prime} (servants of the Inquisitor Prime).
\end{itemize}

\subsection*{The Inquisitor Prime's Hunter}
\index{Inquisitor Prime}\index{patrons!Inquisitor Prime}

Witch hunters who swear to the \textbf{Inquisitor Prime} (see \S\ref{chap:magic}) gain supernatural tools at the cost of Obligation and creeping zealotry. The Prime's Rites are already detailed in the magic chapter; a hunter may access them as a \textbf{Runekeeper} (Thiasos + Codex) or an \textbf{Invoker} (Patron's Symbol). Common choices:

\begin{itemize}
  \item \emph{Rite of the Unclouded Eye} -- pierce illusions and magical deception.
  \item \emph{Rite of the Pure Flame} -- gain +2 dice to resist supernatural influence.
  \item \emph{Rite of the Marked Prey} -- curse a supernatural target, making them easier to track.
\end{itemize}

\textbf{Corruption of the Prime} (summarized): As Obligation grows, the hunter becomes increasingly absolute -- seeing only "pure" and "corrupt", suffering --1 die in morally grey situations. At high tiers, they may gain the \emph{Absolute Judgment} ability, but at the cost of becoming a monomaniacal instrument of the Prime.

\subsection*{Play Style}
\begin{itemize}
  \item Tracks leads, reads scenes, corners quarry with paperwork and steel.
  \item Uses Position and leverage in social spaces before blades are drawn.
  \item Treats magic as both \emph{evidence} and \emph{hazard} (\S\ref{ch:magic}).
  \item May invoke Patron Rites to ward, reveal, or smite -- but each use deepens Obligation.
\end{itemize}

\subsection*{Development Path}
\begin{itemize}
  \item Specialize in \emph{Baneful Techniques} (counter-rites, bindings, exorcism).
  \item Build a network of informants and safe cells (\S\ref{ch:assets-followers}).
  \item Learn one magical tradition deeply to turn it against itself.
  \item For the faithful: advance in the Inquisitor Prime's favor (Embrace Corruption, Purifying Flame talents).
\end{itemize}

\subsection*{Story Hooks}
\begin{itemize}
  \item A vanished ledger ties a noble salon to forbidden rites -- but the salon is protected by a rival inquisitor.
  \item A \emph{sanction} arrives with no seal -- who really wrote it, and what do they want you to ignore?
  \item Your oldest informant begs you to burn a file with your name on it -- it contains evidence that you once used a cursed artefact to save a child.
  \item The Inquisitor Prime appears in a dream: "There is rot in your own order. Find it -- or I will find you."
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white,colframe=gray!70!black,title=Witch Hunter's Eye,fonttitle=\bfseries]
Once per session when confronting occult evidence or a supernatural threat, choose one:
\begin{itemize}
  \item \textbf{Pattern Recognition}: Reduce DV by 1 on the first investigation or Lore roll.
  \item \textbf{Measured Response}: Shift Position one step safer when disrupting active magic or attacking a supernatural creature.
  \item \textbf{Witness Prepared}: Gain +1 die on the first social roll when presenting your authority (writ, badge, or Patron's symbol).
  \item \textbf{Purity's Toll (Inquisitor Prime only)}: Once per session, invoke the Prime's name to reroll a failed resistance roll against a supernatural effect. Mark +1 Obligation.
\end{itemize}
\end{tcolorbox}

\textbf{Sanction Ledger} [4]: Tracks cases where you"ve used authority beyond scope or invoked the Prime's Rites in public. When full, local magical communities become wary -- and the Inquisitor Prime may send a test. Mark a segment when you:
\begin{itemize}
  \item Forge or exceed a writ's limits.
  \item Use a Rite of the Prime in front of witnesses who aren"t sworn allies.
  \item Refuse a demand from your order or patron.
\end{itemize}
Clearing the ledger requires a downtime action: paying fines, doing penance, or destroying evidence of overreach.
