% ============================================================
% Chapter 1: Quick Start
% ============================================================
\chapter{Quick Start}
\label{ch:quickstart}

\section{What You Need to Play}

Fate's Edge is a game where you tell stories together. You need:
\begin{itemize}
  \item \textbf{A Game Master (GM):} The person who runs the world and plays everyone else.
  \item \textbf{Players:} Each of you plays one main character.
  \item \textbf{Dice:} A handful of ten-sided dice (d10s).
  \item \textbf{Character sheets:} To track your hero.
  \item \textbf{This book:} For reference.
\end{itemize}

\section{The Core Loop in 60 Seconds}

Every important action follows this simple loop:

\begin{enumerate}
  \item \textbf{Declare what you do.} Say what you want to achieve and how.
  \item \textbf{The GM sets the stakes.} They tell you the \textbf{Difficulty (DV)} and \textbf{Position}.
  \item \textbf{Roll dice.} Your dice pool is \textbf{Attribute + Skill} d10s.
  \item \textbf{Count successes.} Each 6+ is a success. Each 10 counts as 2 successes.
  \item \textbf{The GM tells you what happens.} Based on your result.
\end{enumerate}

\section{Your Character}

Your character has:
\begin{itemize}
  \item \textbf{Four Attributes:} Body, Wits, Spirit, Presence (rated 1-5).
  \item \textbf{Skills:} Things you're trained in (rated 0-5).
  \item \textbf{Talents:} Special abilities that make you unique.
\end{itemize}

Your dice pool is always \textbf{Attribute + Skill}.

\section{The Dice}

Roll your dice pool. Each die that shows:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Die Result} & \textbf{What It Means} \\
\midrule
6-9 & 1 Success \\
10 & 2 Successes (critical!) \\
1 & 1 Story Beat (SB) for the GM \\
2-5 & Nothing \\
\bottomrule
\end{tabular}
\end{center}

\section{Difficulty (DV)}

The GM sets a target number you need to hit:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{DV} & \textbf{What It Means} \\
\midrule
2 & Routine — almost guaranteed \\
3 & Default — a real challenge \\
4 & Hard — serious resistance \\
5+ & Extreme — a dramatic gamble \\
\bottomrule
\end{tabular}
\end{center}

\section{Position}

Position tells you how risky the action is:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Position} & \textbf{How You Act} & \textbf{Mechanical Effect} \\
\midrule
Dominant & You have the advantage & Re-roll one failure \\
Controlled & Normal risk & No re-rolls \\
Desperate & You're in serious danger & Re-roll one success \\
\bottomrule
\end{tabular}
\end{center}

\section{The Outcome}

Compare your successes to the DV:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Result} & \textbf{Outcome} & \textbf{What Happens} \\
\midrule
S $\ge$ DV, SB=0 & Clean Success & You get what you want, no cost \\
S $\ge$ DV, SB>0 & Success with SB & You get what you want, but the world pushes back \\
0 < S < DV & Partial & You make progress, but not all the way \\
S = 0 & Miss & You fail, and things get worse \\
\bottomrule
\end{tabular}
\end{center}

\section{Boons and Story Beats}

\subsection*{Boons (Your Resource)}
\begin{itemize}
  \item You earn Boons on a \textbf{Partial} (1 Boon) or \textbf{Miss} (2 Boons).
  \item Spend Boons to: re-roll a die, improve your Position, or activate an Asset.
  \item You can hold up to 5 Boons.
\end{itemize}

\subsection*{Story Beats (GM's Resource)}
\begin{itemize}
  \item Every time you roll a 1, the GM gains a Story Beat (SB).
  \item The GM spends SB to add complications, escalate threats, or advance timers.
  \item This keeps the story interesting — 1s are not punishments; they're fuel for drama.
\end{itemize}

\section{Quick Start Example}

\textbf{Lyra} tries to pick a lock while guards patrol.

GM: \textit{``Controlled, DV 3. On a Miss, the guard hears you.''}

Lyra rolls (Wits 3 + Subterfuge 2 = 5 dice): 9, 6, 4, 2, 1

\begin{itemize}
  \item Successes: 9 and 6 = 2 successes vs DV 3
  \item Result: \textbf{Partial} — the lock clicks, but Lyra gains 1 Boon
  \item The 1 gives the GM 1 SB
\end{itemize}

GM: \textit{``The lock opens, but the last pin squeals. I spend the SB: the guard stops and turns his head. You have one beat to hide or act.''}

\section{What's Next?}

This is the core loop. Everything else — magic, combat, advancement, assets — is built on this foundation.

Now you're ready to:
\begin{enumerate}
  \item Create a character (Chapter 2).
  \item Learn the full rules (Chapter 3).
  \item Pick a magical path (Chapters 4-5).
\end{enumerate}

\textit{``The road remembers. Every broken wheel leaves a mark, every lit lamp bears witness. The only question is: what are you willing to owe?''}
