\chapter{World Interaction}
\label{ch:world-interaction}
\index{world interaction}

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Using This Chapter]
This chapter gives you the tools to move through the world, interact with its people, and make your journey meaningful. For detailed descriptions of each region's culture, subgenre, and notable NPCs, see \textbf{Chapter \ref{ch:world-cultures} (World Regions and Cultures)}. The two chapters work together: this one covers \emph{how} to travel and interact; the other tells you \emph{where} and \emph{with whom}.
\end{tcolorbox}

In \textbf{Fate's Edge}, the world is not a backdrop -- it is a partner in the conversation. Dikes groan under black rain in Viterra, clan horns answer across Acasia's ridgelines, Ecktoria's marble halls echo with careful words, and Kahfagia's pilots read storms by taste. Wherever you go, place, culture, and pressure push back.

\section{The Known World: A Geographic Overview}
\label{sec:known-world}
\index{geography}\index{regions}

The following summary orients the major regions relative to one another. (See Chapter \ref{ch:world-cultures} for cultural details.)

\begin{description}
  \item[Ecktoria] \index{Ecktoria} Temperate heartland of the old Utaran Imperium. Borders Acasia to the northeast and Kahfagia to the west. Marble cities, imperial roads, codified law.
  \item[Acasia] \index{Acasia} Lowland of rivers and petty kingdoms, northeast of Ecktoria. Broken province of mercenary companies and border lords. Borders Vhasia to the east.
  \item[Vhasia] \index{Vhasia} Fractured kingdom on a southern peninsula. East of Acasia, west of Viterra. Courtly intrigue, martial tradition, and fractured sun symbolism.
  \item[Viterra] \index{Viterra} Shares the same peninsula with Vhasia, lying further east. Its eastern coast borders a dark inland body (the Blackwater). Hedge‑law, river tolls, and the Last Kingdom.
  \item[Kahfagia] \index{Kahfagia} West of Ecktoria, facing the open ocean. Maritime empire of storm‑flags, deep‑sea pilots, and exotic imports.
  \item[Thepyrgos] \index{Thepyrgos} Occupies the southern part of the great peninsula opposite Vhasia/Viterra. Eastern coast overlooks the Astroegro Straits. City of a Thousand Stairs, center of learning.
  \item[Theona] \index{Theona} Large island off the eastern coast of Viterra. Three isles, causeways, marsh, and the ``No Ninth'' taboo.
  \item[Mistlands] \index{Mistlands} Western shore of the Aberderrian Sea (north of the Amaranthine). Bog, fog, and bell‑lines; protectorate of the Aeler. Direwood lies to the west.
  \item[Linn] \index{Linn} Scattered around the Yloka River and the Aberderrian Sea. Longships reach the Blackwater and raid Theona and Viterra. Fjords, skerries, storm‑oaths.
  \item[Violet Steppe] \index{Violet Steppe} North of the Amaranthine, west of the Aberderrian Sea. Vast grassland, home of the Ykrul nomads.
  \item[Valewood] \index{Valewood} East of the Blackwater. Ancient forest of phasing ruins, star‑roads, and fae courts. Former Lethai empire.
  \item[Aelinnel (Aevrossa)] \index{Aelinnel} South of the Valewood. Gnomish homeland of stone, bough, and bright things. Mist paths, bell‑mounds, tide‑gates.
  \item[Aelaerem] \index{Aelaerem} South of Aevrossa. Halfling downs of gentle valleys and deep barrows. Homeland spoken of only in euphemisms (``The Silent Welcome'').
  \item[Zakov] \index{Zakov} Island in the Blackwater, near the mouth of the Yloka River. Pirate haven, reef passages, salt and serpent.
  \item[Oshiira] \index{Oshiira} Southern grasslands east of the Gulf of Oshiira. Grain confederacy, weigh‑houses, and the Vermilion Corps.
  \item[Ashaan] \index{Ashaan} Eastern desert peninsula, south of the Crimson Valley. Veiled Throne, spirit‑binding, and the Red Desert's creeping glass.
  \item[Sekogo] \index{Sekogo} Jungle realm south of the Crimson Valley and west of Ashaan. Green Council, honey‑orchards, and the Ukwe (Speaking Tree).
  \item[Taharka] \index{Taharka} Highland kingdom of the Monsoon Crown. Coffee terraces, ridge‑watchers, and passes that remember every army.
  \item[Fhara] \index{Fhara} Desert peninsula of oasis courts and water‑clerks. Well‑judges, coffee rituals, and the Sunken City of Qalat al‑Madina.
  \item[Sidhi] \index{Sidhi} Diaspora people of the Brass Coast. Lighthouse courts, freedman's seals, and the green flame of memory.
  \item[Pereshi] \index{Pereshi} Highland fire‑temple nation. Ash Prince's seal, fire‑readers, and the ninth coal that never cools.
  \item[Kuvani] \index{Kuvani} Eastern steppe remount clans. Horse lineages, wind‑songs, and the ghost herd of the salt pan.
  \item[Tulkani] \index{Tulkani} Itinerant storytellers and dream‑carriers. Wagon circles, memory‑cords, and the Dreamer who sleeps beneath the road.
  \item[Ngomebe] \index{Ngomebe} Walking cities of the southern plateau. Keystone chargers, roller‑pits, and the Still Voice of abandoned stone.
  \item[Dhahara (Nine Kingdoms)] \index{Dhahara} Elephant kingdoms of the monsoon lowlands. Peacock Throne, river wars, and the ninth kingdom that was never whole.
\end{description}

\subsection*{The Ways Between}
\label{sec:ways-between}
\index{Ways Between}\index{Obishaal}

Beyond the physical map lie the Ways Between -- the dream‑roads and spirit‑paths of Obishaal, realm of dreams and death. Travel here is measured not in miles but in choices. A wrong turn might lead to yesterday. A right turn might lead to a version of yourself who made a different decision and is still alive to regret it.

\paragraph{Entering the Ways:} Thresholds exist in thin places -- crossroads at midnight, bridges that appear only at low tide, caves behind waterfalls, or simply a mist that settles too thick and stays too long. A character may also enter through ritual, dream, or near‑death experience.

\paragraph{Navigation in Obishaal:} Use the standard travel procedures, but the GM may replace the Spades (place) draw with a \textbf{Virtue Test}. The road judges. Acts of honesty, courage, mercy, or self‑sacrifice advance a personal \textbf{Virtue Timer [3]}; acts of greed, cowardice, cruelty, or despair advance a \textbf{Vice Timer [3]}. When one fills, the character ascends (Virtue) to a higher layer or falls (Vice) to a lower one.

\paragraph{The Thirteenth Stone:} At the deepest layer, pressed into the silt of the oldest dream, lies the Thirteenth Stone -- warm, pulsing, and whispering. Here, Ikasha, She Who Sleeps, dreams of freedom. Her voice offers bargains. Her freedom would remake the world.

\section{Game Structure and Time}
\label{sec:game-structure}
\index{Time}\index{Game structure}

Understanding how time works in Fate's Edge helps you navigate both the mechanical and narrative flow of play.

\subsection*{Basic Units}
\label{sec:time-units}
\begin{description}
\item[\textbf{Scene}] \index{Scene} The basic unit of narrative play, covering a specific situation or conflict (Some Time to Significant Time). Resolves a particular question or challenge.
\item[\textbf{Player Turn (Beat)}] \index{Turn} An individual player's action within a scene: Declare action $\rightarrow$ GM sets position $\rightarrow$ roll $\rightarrow$ resolve outcome $\rightarrow$ manage consequences.
\item[\textbf{Round}] \index{Round} Simultaneous or near‑simultaneous actions within a scene (primarily for combat), representing a few seconds of real time.
\item[\textbf{Session}] \index{Session} One complete game session (typically 3--6 hours), containing 2--4 major scenes and resolving significant narrative progress.
\item[\textbf{Downtime}] \index{Downtime} The narrative time between scenes, used for recovery, advancement, and off‑screen activities. Measured in days, weeks, or months depending on fiction.
\item[\textbf{Campaign}] \index{Campaign} Entire story arc (6--20+ sessions) with major character development and lasting consequences.
\end{description}

\section{Movement and Positioning}
\label{sec:movement-positioning}
\index{Movement}\index{Position}

Space is tracked with \textbf{range bands} and \textbf{Position} (see Chapter \ref{ch:core-mechanics} for full rules).

\subsection*{Range Bands}
\index{Range}
\begin{description}
\item[\textbf{Close}] \index{Range!Close} Touching distance: grapples, knife‑work, hand on a relic.
\item[\textbf{Near}] \index{Range!Near} Same room/yard/deck; a rush away.
\item[\textbf{Far}] \index{Range!Far} Same site but distant; requires route or time to reach.
\item[\textbf{Absent}] \index{Range!Absent} Off‑screen; requires scene change or significant effort to interact.
\end{description}

\subsection*{Movement Actions}
\label{sec:movement-actions}
\begin{itemize}
\item \textbf{Move}: Shift one range band as a \emph{beat}.
\item \textbf{Dash}: Shift two bands as your full action (terrain may require a roll).
\item \textbf{Melee Flag}: Mark when two parties are in Near range and directly engaged in combat.
\end{itemize}

\subsection*{Position States}
\index{Position!Dominant}\index{Position!Controlled}\index{Position!Desperate}
\begin{description}
\item[\textbf{Dominant}] \index{Position!Dominant} You have cover, leverage, or ritual footing. Failure still leaves options.
\item[\textbf{Controlled}] \index{Position!Controlled} Standard case: exposed lanes, rivals near, watchful eyes. Failure has teeth, but not ruin.
\item[\textbf{Desperate}] \index{Position!Desperate} Bad ground, bad odds, bad timing. Failure is severe; success may bring extra XP.
\end{description}

\paragraph{Position Shifting:}
\begin{itemize}
\item GM can spend \textbf{1 SB} to worsen Position by one step.
\item Player can spend \textbf{1 Boon} to improve Position by one step (once per action).
\item Narrative triggers (flanking, reinforcements, etc.) can shift Position without cost.
\end{itemize}

# Travel Framework (Revised)

I've restructured the travel system to address the issues we identified. The key changes:

1. **Travel as Scene Framing** — Each leg now has a dramatic question that must be answered
2. **Character-Driven Complications** — Hazards tie to character backstories, Patrons, and ongoing plot threads
3. **Meaningful Tradeoffs** — Choices have narrative consequences, not just numeric modifiers
4. **Regional Integration** — Travel is inseparable from the region's identity
5. **Patron Engagement** — Travel is when Patrons test, demand, or reward

---

\section{Travel Framework}
\label{sec:travel-framework}
\index{Travel}\index{Travel framework}

Travel in Fate's Edge is not a procedural buffer between scenes—it is where the world asserts its claims. Every journey is a negotiation with distance, weather, and the debts the land remembers. The road itself is a character, and it has opinions.

\subsection*{Core Principle: Travel as Scene, Not Encounter}
\label{sec:travel-core}

A travel leg is not a series of random rolls. It is a **framed scene** with a dramatic question, meaningful choices, and consequences that ripple forward.

Before any travel roll, the GM answers three questions:

\begin{enumerate}
\item \textbf{What is at stake?} What changes if the party succeeds or fails?
\item \textbf{What choice must the party make?} What tradeoff are they being asked to accept?
\item \textbf{How does this connect to the campaign?} What ongoing thread does this leg touch?
\end{enumerate}

If a travel leg has no answer to all three questions, resolve it in a single sentence and move on. Not every mile needs a scene.

\subsection*{Travel Process}
\label{sec:travel-process}

\begin{enumerate}
\item \textbf{Frame the Leg:} Name origin and destination. State the dramatic question. Set the Travel Timer (4--10 segments based on difficulty and urgency).
\item \textbf{Draw Prompts:} Draw one card from each suit from the regional deck. These establish terrain (♠), actors (♥), pressure (♣), and leverage (♦).
\item \textbf{Assign Roles:} Players take travel roles (see below). Each role is a \emph{scene}, not a passive bonus—the player describes what their character does.
\item \textbf{Play the Leg:} Players take actions to advance the timer, resolve complications, or pursue opportunities. Rolls that generate \textbf{1}s give the GM SB to introduce hazards or escalate pressure.
\item \textbf{Resolve:} When the timer fills, the party arrives—but the consequences of their choices arrive with them.
\end{enumerate}

\subsection*{The Dramatic Question}
\label{sec:travel-question}

Every leg must have a clear dramatic question. Examples:

\begin{itemize}
\item \textit{Will the party reach the siege before the supply line is cut?} (Timer: 6, ticking down)
\item \textit{Can they cross the Belworth without alerting the border patrol?} (Timer: 4, risk of Exposure)
\item \textit{Will the Traveler's test break them before they reach the shrine?} (Timer: 8, with Patron complications)
\end{itemize}

The dramatic question frames every roll. Successes advance toward the answer; failures complicate it.

\subsection*{Travel Roles}
\label{sec:travel-roles}

Each role is a \emph{scene role}, not just a passive bonus. Players describe their character's approach and the GM sets Position and DV accordingly.

\begin{description}
\item[Guide] \index{Travel!Guide} Reads the land and sets the course. \textit{What path do you choose, and why?} On success, tick the Travel Timer +1. On a failure, the path reveals a hidden cost (tick a Complication Timer or introduce a hazard).
\item[Scout] \index{Travel!Scout} Moves ahead, reading the terrain and those who move through it. \textit{What do you notice, and what do you do with that knowledge?} On success, the party cannot be surprised on the first encounter of the leg. On a failure, the scout returns with a warning—and something following.
\item[Quartermaster] \index{Travel!Quartermaster} Manages the party's resources and morale. \textit{What do you stretch, and what do you sacrifice?} On success, prevent one Supply loss or recover 1 spent Supply. On a failure, the shortage creates a specific need (water, food, medicine) that must be addressed.
\item[Watch] \index{Travel!Watch} Keeps vigilance against threats seen and unseen. \textit{What do you watch for, and what do you miss?} On success, the party gains +1 Position on the first defense roll of any ambush. On a failure, the GM may spend SB to introduce a threat from the direction the watch did not guard.
\end{description}

\textbf{Role Synergy:} When two or more roles succeed on the same leg, the party gains \textbf{Momentum}. Momentum can be spent to:
\begin{itemize}
\item Tick the Travel Timer +1 additional segment.
\item Reduce the severity of a complication by one step.
\item Gain +1 Position on the first roll of the destination scene.
\end{itemize}

When two or more roles fail, the GM gains \textbf{Setback}. Setback can be spent to:
\begin{itemize}
\item Introduce a complication at the destination.
\item Tick a Threat Timer without spending SB.
\item Increase the DV of the first roll at the destination by 1.
\end{itemize}

\subsection*{The Travel Timer}
\label{sec:travel-timer}

The Travel Timer represents the party's progress through the leg. Default sizes:
\begin{itemize}
\item \textbf{4 segments:} Short leg, routine travel, or urgent transit.
\item \textbf{6 segments:} Standard leg, moderate distance, or tense journey.
\item \textbf{8 segments:} Long leg, difficult terrain, or dangerous passage.
\item \textbf{10+ segments:} Epic journey, hostile environment, or pursuit.
\end{itemize}

\textbf{Ticking the Timer:}
\begin{itemize}
\item \textbf{Success:} Tick +1 (or +2 on a Critical Success).
\item \textbf{Partial:} Tick +1, but the GM gains 1 SB.
\item \textbf{Miss:} No tick; the GM gains 2 SB and may introduce a complication.
\item \textbf{Complication:} The GM may spend SB to tick a Complication Timer instead of or in addition to the Travel Timer.
\end{itemize}

When the Travel Timer fills, the leg is complete. The party arrives—but unresolved complications carry forward to the destination.

\subsection*{Complication Timers}
\label{sec:complication-timers}

Each leg may have one or more Complication Timers representing accumulating pressure. Examples:
\begin{itemize}
\item \textbf{Storm Front [6]:} A storm is approaching. When it fills, the weather turns severe.
\item \textbf{Pursuit [4]:} Someone is following. When it fills, they catch up.
\item \textbf{Hollow Attention [8]:} The party has drawn the Hollow's notice. When it fills, an Oath-Keeper appears.
\end{itemize}

Complication Timers tick when:
\begin{itemize}
\item A travel roll fails.
\item The GM spends SB.
\item A role roll generates \textbf{1}s.
\end{itemize}

When a Complication Timer fills, the promised consequence manifests. The party must address it or carry it to the destination.

\subsection*{Supply as Narrative Resource}
\label{sec:travel-supply}

Supply is not a passive attrition track. It is a resource players spend to shape outcomes.

\begin{itemize}
\item \textbf{Starting Supply:} 6 segments at the beginning of a journey.
\item \textbf{Spending Supply:} Players may spend 1 Supply to:
  \begin{itemize}
  \item Reroll a failed travel roll.
  \item Declare a favorable encounter (a safe campsite, a friendly traveler, a cache of resources).
  \item Avoid a hazard before it triggers (the GM must agree).
  \item Gain +1 Position on a role roll.
  \end{itemize}
\item \textbf{Losing Supply:}
  \begin{itemize}
  \item Quartermaster failure: lose 1 Supply.
  \item Harsh terrain or extreme weather: lose 1 Supply per leg.
  \item GM may spend SB to force a Supply loss (1 SB = 1 Supply).
  \end{itemize}
\item \textbf{Running Out:} When Supply reaches 0, the party is \textbf{Depleted}. All travel rolls are made at Desperate Position until Supply is restored.
\item \textbf{Restoring Supply:}
  \begin{itemize}
  \item Successful foraging (Wits + Survival, DV 3): regain 1 Supply.
  \item Reaching a settlement: regain all Supply.
  \item Trade or barter with travelers: regain 1-2 Supply.
  \end{itemize}
\end{itemize}

\subsection*{Character-Driven Complications}
\label{sec:character-complications}

Travel complications should tie to character backstories, Patrons, and ongoing plot threads. The GM should use the following triggers:

\begin{itemize}
\item \textbf{Patron Obligation:} If a character has Obligation to a Patron, the GM may spend 2 SB to introduce a Patron-related complication during travel.
  \begin{itemize}
  \item \textit{The Traveler tests your resolve. The path splits three ways.}
  \item \textit{The Witness demands you record something you'd rather forget.}
  \item \textit{Malachai offers a shortcut—at a price.}
  \end{itemize}
\item \textbf{Complications:} If a character has a complication (Hunted, Forbidden Knowledge, Cursed Lineage, Oath-Bound), the GM may spend 1 SB to introduce a travel complication related to it.
  \begin{itemize}
  \item \textit{Your hunter has found your trail.}
  \item \textit{The knowledge you carry draws the attention of a local power.}
  \item \textit{Your curse manifests in this cursed place.}
  \end{itemize}
\item \textbf{Bonds:} If a character has a Bond with another PC, the GM may offer a Devil's Bargain during travel: resolve the leg smoothly, but your Bond is strained.
  \begin{itemize}
  \item \textit{You can save time by taking the dangerous path, but your ally will have to carry you if you fall.}
  \item \textit{You can secure supplies, but you'll owe a favor to a faction your Bonded ally distrusts.}
  \end{itemize}
\end{itemize}

\subsection*{Regional Travel Decks}
\label{sec:travel-decks}
\index{Travel!regional decks}

Each region has a themed prompt list. When a travel leg begins, the GM draws one card from each suit from the relevant regional deck.

\begin{itemize}
\item \textbf{♠ Spades (Place):} The terrain or landmark that defines the leg.
\item \textbf{♥ Hearts (Actor):} The people or factions the party encounters.
\item \textbf{♣ Clubs (Pressure):} The hazards, obstacles, or complications.
\item \textbf{♦ Diamonds (Leverage):} The opportunities, shortcuts, or assets.
\end{itemize}

The GM synthesizes these prompts into the leg's fiction. The highest rank sets the Travel Timer size (2-5 = 4, 6-10 = 6, J/Q/K = 8, A = 10).

\textbf{Regional Decks by Region:}
\begin{itemize}
\item \textbf{Aeler:} Stone, breath, underground, bell-lines.
\item \textbf{Mistlands:} Fog, bells, salt, wraiths, ward-lines.
\item \textbf{Vhasia:} Rivers, chevauchée, border patrols, broken crowns.
\item \textbf{Viterra:} Hedges, ferries, boundary stones, legal disputes.
\item \textbf{Acasia:} The Curse, broken roads, ghost bridges, famine.
\item \textbf{Thepyrgos:} Stairs, canals, synod edicts, Chain-Lanterns.
\item \textbf{Kahfagia:} Sea lanes, tides, privateers, mirror-towers.
\item \textbf{Linn:} Fjords, longships, sudden squalls, oath-keepers.
\item \textbf{Ykrul:} Steppe, remounts, white squalls, sky omens.
\item \textbf{Valewood:} Star-roads, phasing ruins, fae geases.
\item \textbf{Aelinnel:} Tide-paths, counting traps, glamour fog.
\item \textbf{The Ways Between:} Thin places, memory loss, the Hollow.
\end{itemize}

\subsection*{Using Assets and Followers During Travel}
\label{sec:travel-assets}
\index{Travel!assets}\index{Travel!followers}

\begin{itemize}
\item \textbf{Assets:} Spend 1 Boon to activate an asset during travel.
  \begin{itemize}
  \item Reveal a hidden path (tick Travel Timer +1).
  \item Call for emergency aid (reduce a Complication Timer by 2).
  \item Secure safe lodging (restore 1 Supply).
  \end{itemize}
\item \textbf{Followers:}
  \begin{itemize}
  \item Assign followers to travel roles. A Cap 3 Scout follower adds +3 to the Scout's roll.
  \item Followers may take independent actions (scout ahead, secure supplies) at the cost of Exposure or Harm.
  \item High-Cap followers (4-5) can solve significant travel problems once per downtime, but generate 1 SB for the party.
  \end{itemize}
\end{itemize}

\subsection*{Travel Complications}
\label{sec:travel-complications}

Common complications, tied to region and circumstances:

\begin{itemize}
\item \textbf{Hazards:} Weather, terrain, wildlife, environmental spirits.
\item \textbf{Social:} Border checks, local politics, cultural misunderstandings.
\item \textbf{Supplies:} Food shortages, equipment failure, contaminated water.
\item \textbf{Pursuit:} Being followed, hunted, or racing against time.
\item \textbf{Thin Places:} Unplanned crossings into the Ways Between.
\item \textbf{Patron Tests:} The Traveler demands a choice. The Witness requires a record. Malachai offers a shortcut.
\end{itemize}

\subsection*{Travel by Region — Quick Reference}
\label{sec:travel-by-region}

\begin{center}
\begin{longtable}{lll}
\toprule
\textbf{Region} & \textbf{Primary Travel Mode} & \textbf{Key Hazard} \\
\midrule
Ecktoria & Imperial roads, river barges & Bureaucratic delays, processions \\
Acasia & Dirt roads, bridge fords & The Curse (roads that loop) \\
Vhasia & River crossings, mountain passes & Border checks, chevauchée raids \\
Viterra & Hedge-lined lanes, ferries & Boundary disputes, legal checks \\
Kahfagia & Sea lanes, coastal routes & Storms, privateers, tide shifts \\
Thepyrgos & Stairs, canals, coastal ships & Chain-Lantern sweeps, thin air \\
Theona & Causeways, coastal boats & Green Host, ninth taboo violations \\
Mistlands & Levees, fog-shrouded tracks & Bell-line failures, wraiths \\
Linn & Longships, fjord passages & Sudden squalls, oath-keepers \\
Violet Steppe & Horseback, remount stations & White squalls, grassfires \\
Valewood & Star-roads, game trails & Phasing ruins, fae geases \\
Aelinnel & Tide-paths, root-galleries & Glamour fog, counting traps \\
Aelaerem & Hedge-tunnels, chalk downs & Hollow attention, missing thresholds \\
Zakov & Reef passages, night boats & Tidal shifts, syndicate feuds \\
Oshiira & Canals, grain barges & Dust storms, weigh-house audits \\
Ashaan & Caravan routes, spirit roads & Veil failures, Black Hand patrols \\
Sekogo & Jungle paths, honey-trails & Thorn-spirits, Ukwe demands \\
Taharka & Mountain passes, monsoon routes & Landslides, thin air, spirit demands \\
Fhara & Oasis routes, desert trails & Dry wells, sandstorms, coffee rituals \\
Sidhi & Coastal ropeways, lighthouse lanes & Customs inspections, reef hazards \\
Pereshi & Fire-temple roads, high passes & Ashfall, name-erasure, bridge recitations \\
Kuvani & Steppe trails, remount strings & Horse sickness, ghost herds \\
Tulkani & Wagon circles, story-roads & Ninth night, Hollow visits \\
Ngomebe & Walking cities, roller-paths & Keystone cracks, root-binder sabotage \\
Dhahara & River barges, elephant routes & Monsoon floods, border disputes \\
\bottomrule
\end{longtable}
\end{center}

\subsection*{GM Quick Cues}
\label{sec:travel-gm-cues}

\begin{itemize}
\item \textbf{Frame Every Leg:} State the dramatic question before any roll.
\item \textbf{Tie to Characters:} Use Patron Obligation, Complications, and Bonds as complication sources.
\item \textbf{Make Tradeoffs Meaningful:} Choices should have narrative consequences, not just numeric modifiers.
\item \textbf{Use Regional Decks:} The region should shape the nature of every encounter.
\item \textbf{Supply as Resource, Not Tax:} Players should choose when to spend Supply, not just watch it drain.
\item \textbf{Momentum and Setback:} Use these to reward coordination and punish disarray.
\item \textbf{Arrival Matters:} The consequences of travel choices should arrive with the party.
\end{itemize}

\subsection*{Example: Travel as Scene}

\noindent \textbf{Setup:} The party is traveling from Vhasia to Viterra across the Belworth River. The GM frames the leg:

\begin{quote}
\textit{The dramatic question: Will you cross the Belworth without alerting the Vhasian border patrol?}

\textit{♠ 7: The Pont-du-Tithe bridge. \textit{♡ J: A Vhasian patrol captain who recognizes one of you.}
\textit{♣ 5: A sudden fog rolls in from the Mistlands.}
\textit{♦ Q: A hidden crossing the ferrymaster knows—for a price.}

\textit{The crossing is watched. The fog could hide you—or betray you. The ferrymaster's price is a secret: he wants to know what happened to his daughter, who disappeared in Viterra a decade ago.}
\end{quote}

\noindent \textbf{The Choices:}
\begin{itemize}
\item \textit{Cross at the bridge:} Risk recognition, but fastest route (Timer: 4).
\item \textit{Use the hidden crossing:} Safer, but requires negotiation and carries a debt (Timer: 6, plus an Obligation to the ferrymaster).
\item \textit{Wait for the fog:} Slow, but grants +1 Position on Stealth rolls (Timer: 8).
\end{itemize}

\noindent \textbf{The Stakes:}
\begin{itemize}
\item If they cross at the bridge and are recognized, they gain Exposure with the Vhasian border patrol (a Complication Timer that follows them to Viterra).
\item If they use the hidden crossing, they owe the ferrymaster—and his daughter's fate is an unresolved plot thread.
\item If they wait for the fog, the patrol may find them anyway (tick a Pursuit Timer).
\end{itemize}

\noindent \textbf{Resolution:} The party chooses the hidden crossing. They negotiate with the ferrymaster, promising to investigate his daughter's disappearance. They cross safely, but they now carry an obligation to a Viterran ghost. The GM notes the debt; it will return when least expected.

\subsection*{Summary}

Travel in Fate's Edge is not a procedural interlude. It is where the world's claims are made visible—where Patrons test, debts compound, and choices shape the destination before the party arrives. Every leg should feel like a scene, not a chore.

\section{Narrative Time}
\label{sec:narrative-time}
\index{Time!narrative}
Time is measured by \emph{importance} rather than duration.
\begin{description}
\item[\textbf{A Moment}] \index{Time!Moment} A glance, a strike, a whisper over a law‑stone.
\item[\textbf{Some Time}] \index{Time!Some Time} A skirmish, a negotiation, a careful climb.
\item[\textbf{Significant Time}] \index{Time!Significant Time} Hours of march, rites, audits, stakeouts.
\item[\textbf{Days}] \index{Time!Days} Drills, recoveries, research, roadwork.
\end{description}

\section{Social Interactions}
\label{sec:social-interactions}
\index{Social}\index{Culture}
Social scenes use the same engine with \textbf{cultural color} (see Chapter \ref{ch:core-mechanics}).

\subsection*{Cultural Skill Emphases}
\label{sec:cultural-skills}
\index{Social!culture}
Different regions value different approaches. Use the table in \textbf{Chapter \ref{ch:world-cultures} (World Regions and Cultures)} to see what skills are most effective where. In general:

\begin{center}
\begin{longtable}{ll}
\toprule
\textbf{Region} & \textbf{Valued Approach} \\
\midrule
Viterra & Rapport with parishes; Sway for markets; Command under writ. \\
Acasia & Rapport for kin‑bridges; Command with banner‑rights; Deceive risks honor timers. \\
Ecktoria & Sway in salons; Deceive at court; Perform in temple fora. \\
Kahfagia & Rapport aboard ship; Sway at piers; Command on a storming deck. \\
Vhasia & Presence for courtly etiquette; Sway for diplomacy; Insight for hidden loyalties. \\
Thepyrgos & Lore for academic debate; Sway for rhetorical argument; Investigation for uncovering truth. \\
Mistlands & Survival for reading fog signs; Resolve for resisting spirit influence; Ritual for bell‑law. \\
Ykrul & Command for leadership; Performance for boast and saga; Survival for steppe navigation. \\
Sidhi & Sway for negotiation; Subterfuge for smuggling; Investigation for lighthouse justice. \\
Tulkani & Performance for storytelling; Insight for reading truth; Survival for road etiquette. \\
\bottomrule
\end{longtable}
\end{center}

\subsection*{Social Stakes \& Timers}
\label{sec:social-timers}
\index{Social!stakes}\index{Timers!social}
Your GM may use visible timers to track social progress:
\begin{itemize}
\item \textbf{Alliance Timer (Viterra):} Parishes and guilds come to your side.
\item \textbf{Honor Timer (Acasia):} Feasts, oaths, wyrd -- trust builds (or frays).
\item \textbf{Bureau Timer (Ecktoria):} Stamps, seals, approvals -- delay is pressure.
\item \textbf{Trust Timer (Kahfagia):} Pilots and crews extend favors and routes.
\item \textbf{Courtesy Timer (Aelinnel):} Fae negotiations -- speak true or lose standing.
\item \textbf{Hollow Attention Timer (Aelaerem/Mistlands):} The uninvited guest draws near.
\end{itemize}

\subsection*{Hospitality as a Social Mechanic}
\label{sec:hospitality}
\index{Social!hospitality}

In many cultures -- especially Aelaerem, Ykrul, Tulkani, Ubral, and Fhara -- hospitality is sacred. Accepting a cup, a meal, or a night's shelter creates a bond. Breaking hospitality is worse than breaking a sworn oath.

\paragraph{Hospitality Bond:} When you accept hospitality (food, drink, or shelter) from a host in a culture that honors guest‑right, you and the host gain a mutual \textbf{Hospitality Bond [3]}. While the bond has segments, neither party may take hostile action against the other. The bond ticks when either party lies to the other, breaks a promise made under the host's roof, or acts to harm the other's interests. When it fills, the bond snaps -- and the violator gains a \textbf{Black Mark} in that culture (−1 die on all social rolls with that culture for one season).

\paragraph{Refusing Hospitality:} Refusing an offered cup, meal, or shelter is a grave insult in hospitality cultures. The GM may spend 1 SB to worsen Position by one step in all social interactions with that host and their community for the remainder of the visit.

\section{Supply and Resources}
\label{sec:supply}
\index{Supplies}\index{Supply Timer }
Track scarcity with a \textbf{Supply Timer } shared by the party's expedition.
\begin{center}
\begin{longtable}{cl}
\toprule
\textbf{Segments} & \textbf{State \& Effects} \\
\midrule
0 (Full) & Well‑provisioned; no penalty. \\
2 (Low) & Minor frictions; --1 to resource checks. \\
3 (Dangerous) & Each PC gains \emph{Fatigue 1}. \\
4 (Empty) & Severe penalties; desperate measures. \\
\bottomrule
\end{longtable}
\end{center}

\subsection*{Foraging and Subsistence by Region}
\label{sec:foraging}
\index{Supplies!foraging}

Different regions offer different opportunities (and risks) for finding food and water.

\begin{center}
\begin{longtable}{lll}
\toprule
\textbf{Region} & \textbf{Forage DV} & \textbf{Typical Finds} \\
\midrule
Forest (Valewood, Aelinnel) & 2--3 & Berries, mushrooms, roots, small game \\
Steppe (Ykrul, Kuvani) & 3--4 & Edible roots, tubers, small game \\
Desert (Ashaan, Fhara) & 4--5 & Cactus water, insects, rare desert herbs \\
Marsh (Mistlands, Theona) & 3--4 & Cattails, frogs, medicinal mud \\
Mountain (Ubral, Pereshi) & 3--4 & Mountain herbs, small game, clean springs \\
Coast (Linn, Sidhi, Zakov) & 2--3 & Shellfish, seaweed, fish \\
Tundra (Northern Violet Steppe) & 4--5 & Lichens, frozen berries, small game \\
\bottomrule
\end{longtable}
\end{center}

\section{Engaging the World -- Player Actions}
\label{sec:player-actions-world}
\index{Procedures!world}
\begin{itemize}
\item \textbf{Scout \& Signal:} A follower can make the next travel action \emph{Dominant} (mark Exposure or Harm 1 on them -- see Chapter \ref{ch:assets-followers}).
\item \textbf{Local Color:} Briefly state what locals notice about you; GM offers a small fictional edge \emph{or} a tempting timer -- choose.
\item \textbf{Mark the Map:} On arrival, declare one change to the fiction (new ford, patron's shrine, toll‑skip). GM may attach a minor timer as cost.
\item \textbf{Asset Activation:} Spend 1 Boon to activate an asset dramatically during a scene (see Chapter \ref{ch:assets-followers}).
\item \textbf{Follower Assistance:} Have a follower assist your actions for bonus dice (max +3 from all sources, see Chapter \ref{ch:core-mechanics}).
\item \textbf{Invoke Cultural Knowledge:} If your character has a relevant background or has spent significant time in a region, you may ask the GM one straightforward yes/no question about local customs, taboos, or safe practices.
\end{itemize}

\subsection*{World Interaction -- Quick Reference}
\label{sec:world-interaction-quickref}

\begin{center}
\begin{longtable}{ll}
\toprule
\textbf{When you want to...} & \textbf{Use this approach} \\
\midrule
Travel quickly & Take the Guide role, prioritize Speed intent \\
Travel safely & Take the Scout role, prioritize Stealth or Survey intent \\
Learn local customs & Ask a local, invoke cultural knowledge, observe before acting \\
Gain a patron & Complete a favor, offer a service, impress with skill or honor \\
Avoid attention & Travel with Stealth intent, keep to back roads, pay bribes \\
Negotiate a dispute & Invoke common ground, offer a compromise, use social skills \\
Survive the wilds & Take the Quartermaster role, forage, find shelter \\
Honor hospitality & Accept offered food/drink, share a story, never draw first blood \\
Cross a thin place & Offer a memory, speak a true name, leave an offering \\
\bottomrule
\end{longtable}
\end{center}

\section{Summary}
\label{sec:world-summary}
\index{World interaction!summary}
The world has opinions. Movement is timers and color, position rises and sinks with weather and words, and every suit you draw speaks in a regional accent. Ask the land for a favor -- then pay it back on the road.

\paragraph{Remember:} Every interaction with the world is an opportunity. Use your assets, deploy your followers, and engage with the setting actively. The world responds to your choices, and every journey changes both you and the places you pass through. The road is not a ledger -- it is a conversation. Speak carefully. Listen closely. And always leave something behind for the next traveler.
\end{tcolorbox}