\chapter{World Powers and Influence}
\label{ch:world-powers-obligation}
\index{terrestrial patrons}\index{influence}\index{factions}

\section{Of Patrons, Runes, and Invokers}
\label{sec:invoker-lore}

\begin{quote}
``You wish to walk the road of power? Then listen well. The world is old, and older still are the voices beneath it. We call them \textit{Patrons}, though they were never sworn to us. They are the tides that move unseen, the keepers of forgotten bargains, the sleepers beneath the stone and the stars. To call upon them is to dip a hand into a river that has carved mountains.''
\end{quote}

\subsection{The Patrons}
\index{Patrons}

Patrons are vast intelligences---not gods, though some worship them as such. They are embodiments of \textit{concepts} and \textit{forces} rather than sovereigns. Each offers power, but always with cost: fatigue, scars upon fate, or a slow unweaving of one's own story.

To entreat a Patron is to risk being marked. Their Rites are gifts and snares both.

\begin{tcolorbox}[
  colback=gray!5!white,
  colframe=gray!60!black,
  enhanced,
  title=\textbf{Iconic Ascension: The Fate of Great Power},
  fonttitle=\bfseries,
  left=1em,
  right=1em,
  top=0.75em,
  bottom=0.75em,
  borderline={0.5pt}{0pt}{gray!60!black},
]
\textbf{In Fate's Edge, power has its price.}  
When a being rises beyond mortal scale, its identity dissolves and reforms into something larger, older, and less free. This transformation is known as \textbf{Iconic Ascension}.

\medskip
\textbf{Icons are not individuals.} They are \emph{forces}---living embodiments of desires, patterns, and long-term agendas. Their power is immense, but their agency is small. An Icon cannot freely choose its actions; its nature compels it.

\medskip
\textbf{Icons cannot touch the world directly.} Their presence is too vast, their will too diffuse. Instead, they act through intermediaries:
\begin{itemize}[noitemsep,topsep=2pt]
    \item \textbf{Chosen} who carry fragments of their purpose
    \item \textbf{Devotees} who interpret their omens
    \item \textbf{Agents} who benefit from their favor
\end{itemize}

\medskip
Icons shape events not with hands, but through \textbf{boons}, \textbf{omens}, and \textbf{shifts in fate} granted to those who serve their desires---whether fleeting hungers or ancient ambitions woven into the world's fabric.

\medskip
\textbf{To serve an Icon is to wield a living idea.} Their gifts are potent, but their influence reshapes those who bear them. Their will echoes in mortal choices long after their voices fade.
\end{tcolorbox}

\subsection{Choosing a Patron}
\label{subsec:choosing-a-patron}

Not every Patron walks in every story. The following pages present a curated selection of powers. \textbf{You are not expected to read them all before creating a character.}

\medskip
For a first character or a campaign that is just finding its feet, we recommend these \textbf{eight} Patrons. They are mechanically straightforward, thematically flexible, and welcome in almost any game:

\begin{itemize}
  \item \textbf{Oath of Flame \& Light} -- vows, radiance, and righteous smiting. A classic paladin or cleric.
  \item \textbf{The Traveler} -- roads, wayfinding, and safe passage. Low risk, high utility.
  \item \textbf{Inaea (Angel of the Spider)} -- mercy, connection, and the threads that bind families and communities.
  \item \textbf{The Witness} -- truth, memory, and uncomfortable revelations. Perfect for investigators or notaries.
  \item \textbf{Carrion King} -- decay, renewal, and the wisdom of endings. Good for those who like cycles of death and rebirth.
  \item \textbf{Ikasha (She Who Sleeps)} -- shadow, secrets, and latent potential. Ideal for rogues, spies, and those who walk unseen.
  \item \textbf{Grimmir (The Wild Speaker)} -- seasons, sacrifice, and the old pact between field and forest. For druids, rangers, and those who respect the wild.
  \item \textbf{Palinode (Queen of Encores)} -- performance, rapture, and the unfinished song. Perfect for bards, dancers, and anyone who lives for the encore.
\end{itemize}

\medskip
If one of these speaks to you, stop here. Read its entry. Make your character. The other Patrons (in the Appendix) will still be waiting when you are ready for a second campaign.

\medskip
\noindent\textbf{Important:} Your Game Master has the final say on which Patrons exist in their world. Some Patrons are tied to specific cultures or story arcs; others may be dormant, forgotten, or simply not part of the campaign. \textit{Always clear your Patron choice with your GM before you write it on your sheet.}

\medskip
Now, if you are determined to explore the full menagerie---or if your GM has already pointed you toward a particular power---read on. The road is long, and the old voices are patient.

% --- Curated selection of 8 recommended patrons (alphabetical order) ---
\clearpage
\input{patrons/carrion-king}
\clearpage
\input{patrons/grimmir}
\clearpage
\input{patrons/ikasha}
\clearpage
\input{patrons/inaea}
\clearpage
\input{patrons/oath-flame-light}
\clearpage
\input{patrons/palinode}
\clearpage
\input{patrons/traveler}
\clearpage
\input{patrons/witness}
\clearpage

% --- Patron Gifts Table (eight recommended) ---
\renewcommand{\arraystretch}{1.25}
\setlength{\LTpre}{0.5em}
\setlength{\LTpost}{0.5em}

\begin{longtable}{@{}p{4.2cm}p{3.2cm}p{7.2cm}@{}}
\toprule
\textbf{Patron} & \textbf{Thematic Skill (+1)} & \textbf{Gift / Lore Bestowal} \\
\midrule
\endfirsthead

\toprule
\textbf{Patron} & \textbf{Thematic Skill (+1)} & \textbf{Gift / Lore Bestowal} \\
\midrule
\endhead

\midrule
\multicolumn{3}{r}{\emph{Continued on next page}} \\
\bottomrule
\endfoot

\bottomrule
\endlastfoot

Carrion King & Survival &
\emph{Grants the feast of decay, where what is dead becomes seed for what yet lives.}\footnote{The Patron's Gift (Imbuement) costs +1 Influence when invoked (or +1 additional if Pushed).} \\

Grimmir & Survival &
\emph{Grants the wisdom of root and stone, speaking the tongues of leaf and fang.}\footnotemark[\value{footnote}] \\

Ikasha & Stealth &
\emph{Grants the hush between footsteps and the raven's omen at every threshold.}\footnotemark[\value{footnote}] \\

Inaea & Persuade &
\emph{Grants the hearth's warmth---shelter for the weary and mercy for the lost.}\footnotemark[\value{footnote}] \\

Oath of Flame \& Light & Command &
\emph{Grants the fire of dawn, a vow that shields the faithful and sears the faithless.}\footnotemark[\value{footnote}] \\

Palinode & Performance &
\emph{Grants the unfinished cadence: a song that never ends, a mask that never slips, an encore that binds the crowd.}\footnotemark[\value{footnote}] \\

The Traveler & Notice &
\emph{Grants the open way, a restless compass and roads where none were marked.}\footnotemark[\value{footnote}] \\

The Witness & Notice &
\emph{Grants the unblinking gaze that unmasks deceit and remembers every oath.}\footnotemark[\value{footnote}] \\

\end{longtable}

\captionof{table}{Patron Gifts and Thematic Skills (curated selection). The +1 bonus applies only when the fiction aligns with the Patron's domain.}
\label{tab:gift-thematic-map-curated}

% --- Patron Fear Influence Table (eight recommended) ---
\subsection{Patron Fear Influence Table (Curated)}
\label{subsec:patron-fear-table-curated}

The following table summarizes how the curated patrons modify fear effects.

\begin{longtable}{@{}p{3.2cm}p{4.5cm}p{2.8cm}p{5.9cm}@{}}
\toprule
\textbf{Patron} & \textbf{Domain Tags} & \textbf{Influence Type} & \textbf{Fear Effect Modifications} \\
\midrule
\endfirsthead
\multicolumn{4}{c}{{\bfseries Continued from previous page}} \\
\toprule
\textbf{Patron} & \textbf{Domain Tags} & \textbf{Influence Type} & \textbf{Fear Effect Modifications} \\
\midrule
\endhead
\midrule
\multicolumn{4}{r}{{Continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot

Carrion King & Decay, Feast, Scavenging & Reroll & 
Fear tagged \emph{Decay, Hunger, Feast} mutates. On 1 (Fight), frenzy devolves into cannibalistic hunger. \\

Grimmir & Seasons, Wild, Sacrifice & Reroll & 
Fear tagged \emph{Wild, Nature, Season} becomes primal dread. On 6 (Submit), victim freezes as if rooted. \\

Ikasha & Shadow, Whispers, Latent Potential & Amplify & 
Fear tagged \emph{Shadow, Voice, Hidden} deepens. On 3 (Freeze), victim may be paralyzed by whispered secrets. \\

Inaea & Mercy, Hearth, Connection & Reroll & 
Fear tagged \emph{Home, Family, Comfort} twists to terror of loss. On 4 (Flee), victim abandons loved ones ``for their own good.'' \\

Oath of Flame \& Light & Dawn, Vows, Purity & Resonance & 
Fear tagged \emph{Light, Sacred, Vow} also imposes Fatigue. Iconic light turns terror into bodily strain. \\

Palinode & Performance, Rapture, Encore & Amplify & 
Fear tagged \emph{Stage, Crowd, Song} becomes paralyzing stage fright. On 6 (Submit), victim is compelled to perform a humiliating act. \\

The Traveler & Ways, Journeys, Crossroads & Resonance & 
Fear tagged \emph{Lost, Path, Journey} also inflicts Complication Points. Terror of losing the way becomes mechanical hindrance. \\

The Witness & Truth, Revelation & Amplify & 
Fear tagged \emph{Truth, Knowledge, Revelation} intensifies. On 6 (Submit), target may confess dark secrets under perceived Iconic judgment. \\

\end{longtable}


\begin{tcolorbox}[colback=black!3!white, colframe=black!40!white, title=Terrestrial Patrons — Mortal Powers That Collect]
  \label{box:terrestrial-patrons}
  \index{terrestrial patrons}
  
  Not all power flows from ancient Patrons and forgotten bargains. Some power sits in counting-houses, guild halls, and the back rooms of taverns. These are \textbf{Terrestrial Patrons} — mortal factions, noble houses, crime syndicates, or merchant leagues that offer tangible aid in exchange for \textbf{Influence} (the same mechanic as Obligation, tracked on a per-patron basis).
  
  \paragraph{How Terrestrial Patrons Work}
  When you accept a favor from a Terrestrial Patron (access, resources, protection, an Asset), you mark \textbf{+1 Influence} to that patron. Your Influence capacity is the same as Obligation: \textbf{Spirit + Presence}. Exceeding capacity causes Fatigue (1 per segment over), and you cannot call on that patron for further favors until you clear segments.
  
  Clearing Influence requires performing tasks for the patron. The GM sets the scope:
  \begin{itemize}
      \item \textbf{Minor task} (deliver a message, pay a bribe): clears 1 Influence.
      \item \textbf{Significant task} (escort a caravan, break a siege): clears 2 Influence.
      \item \textbf{Major task} (assassinate a rival, broker a treaty): clears 3+ Influence.
  \end{itemize}
  
  \paragraph{Asset Slots}
  A Terrestrial Patron can grant Assets (see \ref{ch:assets-followers}) in exchange for Influence, not XP. The Influence cost is based on the Asset's value and the patron's power:
  
  \begin{center}
  \begin{tabular}{lll}
  \toprule
  \textbf{Asset Type} & \textbf{Minor Patron (Local)} & \textbf{Major Patron (Regional)} \\
  \midrule
  Minor Asset (4 XP value) & 2 Influence & 2 Influence \\
  Standard Asset (8 XP value) & 4 Influence & 3 Influence \\
  Major Asset (12 XP value) & — & 4 Influence \\
  Follower (Cap 2) & 2 Influence & 2 Influence \\
  Follower (Cap 3) & 4 Influence & 3 Influence \\
  \bottomrule
  \end{tabular}
  \end{center}
  
  \noindent \textit{A Minor Patron cannot grant Major Assets — their reach is too short.}
  
  \paragraph{Examples of Terrestrial Patrons}
  \begin{itemize}
      \item \textbf{House Contarini (Vilikari)} — Archivolt access, forged charters, legal immunity.
      \item \textbf{The Ragman (Silkstrand)} — Smuggling routes, night-boat passage, stolen goods.
      \item \textbf{The Morosini (Vilikari)} — Letters of credit, market access, debt collection.
      \item \textbf{The Black Banners (Mercenary)} — Condotta contracts, war mounts, safe passage.
      \item \textbf{The Salt Syndicate (Fhara)} — Water rights, caravan escorts, desert guides.
  \end{itemize}
  
  \paragraph{Why Use Terrestrial Patrons?}
  Terrestrial Patrons allow players to gain Assets and favors without spending XP — but the cost is narrative. A favor from House Contarini may come due in the middle of a heist. A loan from the Morosini may be called when you are penniless and far from home. The Influence track ensures that these relationships have mechanical weight without replacing XP advancement.
  
  For GMs: Terrestrial Patrons are excellent sources of adventure hooks. When a party owes Influence to three different patrons, those patrons will eventually compete for the party's loyalty — and that competition is a campaign.
  
  \end{tcolorbox}
  
\section{Witchcraft, Cults, and Religions}
\label{sec:witchcraft-cults-religions}

The Patrons do not demand worship. They do not build temples, compose scripture, or care which name you speak when you pray. They notice only one thing: \textbf{Influence} (sometimes called \textit{Weight}). A vow kept, a promise honored, a threshold crossed with witness—these are the currencies that move the old powers. Yet mortals are creatures of meaning, not merely transaction. We build cathedrals to flames we do not understand. We whisper confessions into bells that cannot answer. We gather in secret circles to sing the name of a Patron we have never seen, hoping that if we sing loudly enough, someone will listen.

This section describes the \textbf{human (and near-human) faces} of power in the Amaranthine: the religions that shape daily life, the cults that form around specific Patrons or living conduits, and the quiet, patient craft of \textbf{witchcraft}—the art of working at thresholds, tying knots, and reminding the world of promises it has forgotten.

\textbf{Religions} are syncretic traditions—the Everflame, the Temple of Light, the Covenant of Mykkiel—that venerate a single divine principle while drawing power from multiple Patrons. They are the faith of the faithful: public, legal, and often entangled with the state.

\textbf{Cults} are concentrations of purpose: fellowships that form around a specific Patron, a specific interpretation, or a specific living voice—a Cantor whose song has drawn the desperate, an Invoker whose bargains have begun to outrun them. Cults are the faith of the desperate: private, dangerous, and always watched.

\textbf{Witchcraft} is not a faith at all. It is a \textit{method}: the art of listening to Echo, steadying the Veil, and leaning into Flow. Witches work at thresholds—doorways, wakes, crossroads, bedsides—and their power is indistinguishable from skilled attention. Witchcraft is the craft of the cunning: patient, subtle, and older than any temple.

Whether you kneel in a basilica, whisper in a root cellar, or simply tie a red thread around a gatepost, the same law applies: \textbf{the world remembers. And everything you do carries weight.}

\subsection{Religions of the Amaranthine}
\label{sec:religions}

The great religions of the known world are not rivals to the Patrons; they are the masks the Patrons wear when mortals need a single face to pray to. Each religion is a syncretic tradition that venerates a divine principle—the Flame, the Light, the Covenant, the Twin Fires—while drawing power from multiple Patrons. The Patrons themselves do not care about theology; they care about \textit{influence}, about the weight of promises, about the ledger of honor. A priest who invokes the Inquisitor Prime and a paladin who calls upon the Oath of Flame \& Light may believe they serve the same god. They are not wrong – but the Patrons are not gods. They are presences, and they will respond regardless of the name on the prayer.

What follows are the four major religious traditions of the Amaranthine. Each is a web of \textbf{Orders}, each Order devoted to a specific Patron while subsuming that Patron into a larger, unified faith. Each is also a \textbf{Terrestrial Patron} (see \S\ref{sec:terrestrial-patrons}) in its own right: faithful who dedicate themselves to a religion gain access to certain Rites (as Low Rites) and bear \textit{influence} with the religion's hierarchy, not directly to the Patrons. The religion mediates the weight of the covenant.

% ------------------------------------------------
\subsubsection{The Everflame (Ecktoria)}
\label{sec:everflame}

The Everflame is the state religion of Ecktoria, a faith of fire, law, and imperial legitimacy. Its great basilicas burn with perpetual flames; its priests carry censers and gavels. The Everflame teaches that \textit{truth and measure are one}: a lie is a theft, a broken oath is a scorch on the cosmic record. The flame itself is the witness – it remembers every vow sworn before it, and it does not forgive.

\paragraph{The Three Orders of the Everflame}
The Everflame is composed of three great Orders, each devoted to a specific Patron but unified under the Everflame's divine fire.

\begin{itemize}
  \item \textbf{Order of the Oath of Flame \& Light (Orthodox)} – The mainstream clergy and knightly orders. They kindle the sacred flames, perform marriages and coronations, and lead the faithful in procession. Their Rites are those of vows, radiance, and righteous smiting. They are the public face of the Everflame.
  \item \textbf{Order of the Inquisitor Prime (Prosecutorial)} – The hunters of heresy and corruption. They investigate unsanctioned magic, root out demonic cults, and enforce the purity of the flame. In Ecktoria they are called the \textit{Chain-Lanterns} (though that name properly belongs to the Temple of Light's inquisitors; the Everflame uses \textit{Flame-Wardens}). Their Rites are those of judgment, purification, and burning censure.
  \item \textbf{Order of the Oath of Mercy and Grace (Hospitalier)} – The healers and confessors. They run hospices, tend to the sick, and hear the confessions of the dying. Their Rites are those of healing, forgiveness, and the spiral of care. In Ecktoria they are often at odds with the Inquisitor Prime – mercy can look like leniency.
\end{itemize}

\paragraph{Beliefs and Practices}
The Everflame teaches that the world was kindled by a single flame, which then divided into three aspects: the \textit{Gleaming} (justice, oaths, protection), the \textit{Searing} (purification, judgment, truth), and the \textit{Warming} (mercy, healing, forgiveness). Worship involves lighting candles, speaking vows aloud, and making offerings of oil, incense, or written confessions. The most sacred rite is the \textit{Fire Trial}, where a person accused of heresy must hold a flame; if it burns, they are guilty; if it warms, they are innocent. (The Inquisitor Prime's Order controls the flame.)

\paragraph{Where It Is Found}
The Everflame is dominant in Ecktoria and its borderlands. It has a small presence in Kahfagia (where it competes with the lantern-law) and among some Vilikari communities. The Church of the Everflame is deeply tied to the Ecktorian state; to be a citizen is to be of the Flame.

% ------------------------------------------------
\subsubsection{The Temple of Light}
\label{sec:temple-of-light}

The Temple of Light arose in Viterra, born from the hedge‑law kingdom's need for a faith that could unify its patchwork of parishes without the imperial baggage of the Everflame. Where the Everflame is fiery and juridical, the Temple of Light is \textit{pastoral and evangelistic}, but with a stern, magiphobic edge. Its priests are teachers, its knights are protectors of pilgrims, and its missionaries travel as far as the Linn isles and the Thepyrgosian countryside. The Temple distrusts unsanctioned magic; it teaches that arcane power must be witnessed, measured, and licensed – or it is a sin against the Light's clarity.

\paragraph{The Three Orders of the Temple}
The Temple of Light venerates the same three Patrons as the Everflame, but with different emphasis and doctrine. Its inquisitorial branch is the \textbf{Chain-Lanterns}.

\begin{itemize}
  \item \textbf{Order of the Oath of Flame \& Light (Orthodox)} – The core of the Temple. Its members are parish priests, wandering preachers, and the \textit{Dawn-Knights} who guard the roads. They emphasize education, hospitality, and the protection of the innocent – and they are the first to report unlicensed magic.
  \item \textbf{Order of the Inquisitor Prime (Prosecutorial)} – The \textbf{Chain-Lanterns}. A zealous, well‑funded order that operates from fortified chapter houses. They carry brass censers and bells, and they are authorized to test any suspected witch with candle‑smoke, bell‑toll, or the ordeal of the white mirror. In Thepyrgos they are a state‑within‑a‑state, conscripting children with arcane talent into ``rehabilitation.'' In Viterra they are more restrained, but still feared.
  \item \textbf{Order of the Oath of Mercy and Grace (Hospitalier)} – The healers and confessors of the Temple. They run hospices, orphanages, and free clinics. In Viterra they are often the only healers a village can afford. Their Rites are those of unburdening and care – though they never heal a curse without reporting it.
\end{itemize}

\paragraph{Beliefs and Practices}
The Temple teaches that Light is not a flame but a \textit{lens}. The divine shines through all things, but it is refracted into three colors: Gold (justice and protection), Silver (truth and correction), and White (mercy and healing). Worship is quiet: morning prayers, lighting a single candle, reading from the \textit{Codex of Illumination} (a collection of parables and legal maxims). The Temple is evangelistic – it sends missionaries to the Linn, the Mistlands, and even into the Thepyrgosian countryside, where its simple message of mercy contrasts with the Chain-Lanterns' severity (though the missionaries and the inquisitors work from the same budget).

\paragraph{Where It Is Found}
The Temple has its strongest foothold in Viterra, but it is also the dominant faith in large parts of Thepyrgos (especially among the common folk, as opposed to the synod's intellectual elite). It has a quiet, growing presence among the Linn – a ``crusade of education'' that builds schools and hospitals, not cathedrals.

% ------------------------------------------------
\subsubsection{The Covenant (of Mykkiel)}
\label{sec:covenant}

The Covenant is the youngest of the three great religions, born from the ashes of the Sidhi homeland; the Law that they carried with them in their travels. It is \textit{legalistic, stern, and uncompromising}. Its God is the Law – not human law, but the divine order that underpins reality. The Covenant holds that the universe is a contract between the divine and the mortal. To break an oath is to tear the fabric of the world; to keep it is to strengthen the Covenant. The Patron Mykkiel (called the Arbiter) is the living seal of this contract.

\paragraph{The Three Orders of the Covenant}
Unlike the Everflame and the Temple, the Covenant's primary Patron is Mykkiel, the Arbiter of the Covenant. The other two Patrons serve as supporting offices.

\begin{itemize}
  \item \textbf{Order of Mykkiel (Orthodox)} – The judges, notaries, and law‑scholars. They settle disputes, draft contracts, and administer the Covenant Courts. Their Rites are those of binding oaths, weighing testimony, and sealing judgments. They are the public face of the Covenant.
  \item \textbf{Order of the Inquisitor Prime (Prosecutorial)} – In the Covenant, the Inquisitor Prime is the \textit{Accuser} – the one who brings hidden sins to light. They are not hunters of heretics but \textit{auditors of the soul}. They investigate broken oaths, expose secret perjury, and demand restitution. Their Rites are those of revelation and correction. They carry no swords – only iron scales and a copy of the \textit{Canons of Obligation}.
  \item \textbf{Order of the Oath of Flame \& Light (Hospitalier – Healing Focus)} – A surprising inclusion: in the Covenant, the Oath of Flame \& Light is not a warrior's path but a \textit{healer's}. They tend to the wounded in body and soul, but their healing always comes with a \textit{contract}: the patient must swear to live a more lawful life. Their Rites are those of purification, cleansing, and bound mercy.
\end{itemize}

\paragraph{Beliefs and Practices}
The Covenant teaches that every act of kindness, every kept promise, every honest word strengthens the divine contract; every lie, every broken oath, every injustice weakens it. The Covenant's symbol is a \textit{broken chain} – signifying that even when empires fall, the law remains. Worship is conducted in courts, not temples; a service involves reading a case, hearing testimony, and pronouncing a judgment. The most devout often become itinerant judges, bringing the Covenant's justice to villages that have never seen a magistrate.

\paragraph{The Covenant in Ashaan – A Secret Resistance}
In the magocracy of Ashaan, where the Veiled Sisters rule through terror and spirit‑binding, the Covenant is forbidden. Yet it survives in secret. The \textit{Breath‑less} resistance network has adopted the Covenant's teachings as a moral framework: they believe that the Veiled Sisters are oath‑breakers of the highest order, and that a new Covenant – a just law – will one day be established over the ashes of the Sisters' tyranny. In the catacombs of Galanina, Covenant judges preside over \textit{trials in absentia} of the Sisters, sentencing them to nothing less than the restoration of order. The Patron Mykkiel does not care whether the accused are present – the weight of the broken oath remains regardless.

\paragraph{Where It Is Found}
The Covenant is strongest in Galanina (where it blends with Pereshi legal traditions), among the Fhara ruling class, and in the secret cells of Ashaan. It has a small but growing presence in Thepyrgos, where synod clerks find its clarity appealing. In Ecktoria, it is viewed as heretical (the Everflame claims sole authority over divine law). In Vhasia, it is tolerated as a philosophy but not as a church.

% ------------------------------------------------
\subsubsection{The Pereshi Faith – The Twin Flames}
\label{sec:twin-flames}

In the eastern provinces and among the Pereshi diaspora, a different light burns: the \textit{Twin Flames}. This faith teaches that the universe is a struggle between two primordial fires – the \textit{Pure Flame} of truth, order, and life, and the \textit{Consuming Flame} of deceit, chaos, and death. Unlike the Everflame's single fire, the Pereshi hold that both fires are sacred; the duty of the faithful is to feed the Pure Flame and starve the Consuming one through acts of honesty, purification, and the preservation of knowledge. The Twin Flames have no single god, but rather a host of \textit{radiant spirits} (which outsiders recognise as Patrons such as the Oath of Flame \& Light, the Witness, and the Pale Shepherd). The religion is intensely legalistic about ritual purity, but it lacks a central hierarchy; each temple is autonomous, governed by a \textit{Flame-Keeper}.

\paragraph{The Two Orders (loose fellowships)}
\begin{itemize}
  \item \textbf{Keepers of the Pure Flame} – Priests, healers, and archivists. They tend the eternal lamps in fire‑temples, copy holy texts, and perform ceremonies of purification. They invoke the Oath of Flame \& Light for blessings and the Witness for truthful testimony.
  \item \textbf{Watchers Against the Maw} – A martial order devoted to hunting undead, demons, and unnatural corruption. They are the ones who burn the bodies of those touched by the Consuming Flame. They invoke the Pale Shepherd to guide spirits and the Inquisitor Prime to expose hidden defilement.
\end{itemize}

\paragraph{Beliefs and Practices}
The Twin Flames teach that every person carries a small spark of the Pure Flame. Through good deeds, honest speech, and proper ritual cleanliness, that spark grows; through lies, cruelty, and contact with the dead (unless ritually purified), it dims. Fire‑temples are open to the sky, with a central basin of sand where the eternal flame burns. Worship consists of pouring offerings of oil or clarified butter into the flame while reciting the \textit{Twenty‑One Names of Radiance} (a list of virtues). The dead are neither buried nor cremated; they are placed in \textit{Towers of Silence} – open‑topped stone cylinders – where the elements and carrion birds cleanse the bones. The bones are then ground and used to make ceremonial chalk for temple circles.

\paragraph{Where It Is Found}
The Twin Flames are the majority faith in the Pereshi highlands (the region between Thepyrgos and the eastern steppe) and among Pereshi communities in Zakov, Kahfagia, and the Dolmis ports. In Ecktoria they are tolerated as a ``philosophical sect'' of the Everflame. In Viterra they are viewed with suspicion because of their dealings with the dead, but the Temple of Light has never actively persecuted them – their distrust of sorcery aligns.

% ------------------------------------------------
\paragraph{Religious Conflict and Cooperation}
The four religions do not wage holy war against each other – not openly. Their priests argue theology; their inquisitors watch each other; their missionaries compete for converts. But they also cooperate. A Covenant judge might consult an Everflame priest on a case involving a sanctified oath; a Temple healer might staff a Covenant hospice; a Pereshi Flame‑Keeper might share fire‑lighting duties at a joint ceremony. The Patrons themselves are indifferent to human sectarianism; they accept influence from any and all.

What the faithful share is a belief that \textit{law is sacred} and that \textit{witness is binding}. Whether the flame, the light, the covenant, or the twin fires, the divine is that which holds you accountable. And the Patrons – whatever names they are called – are always watching.

% ------------------------------------------------
\subsubsection{Other Traditions and Cults}
\label{sec:other-religions}

Beyond the four great syncretic faiths, the Amaranthine harbors a host of local traditions, mystery cults, and folk practices. These are not organised religions in the sense of the Everflame or the Covenant; they lack central hierarchies, formal clergy, or unified doctrine. Yet they shape the spiritual landscape of their regions as powerfully as any basilica or fire-temple. Most are tolerated (or ignored) by the great faiths, though some are actively suppressed.

\paragraph{The Grey Pilgrims (The Traveler)}
A wandering order of Lethai‑ar and mixed‑blood devotees who maintain the old way‑cords of the Valewood and the star‑roads of the eastern marches. They wear grey cloaks woven from mist and speak only in riddles. Their creed is simple: \textit{``The road does not end. It only turns.''} They accept no payment for their guidance, only a lock of hair, which they knot into a way‑cord. A Grey Pilgrim is never lost, but they are also never home. Found in the Valewood, the Linn coast, and the passes of the Aelerian mountains.

\paragraph{The Daughters of the Cord (The Bitter Root)}
A secretive, all‑female cult that traces its origin to a renegade Sister of the Pale Shepherd. They believe that all obligations are threads that can be cut – and that the highest form of mercy is to sever a promise without asking permission. They run hospices and soup kitchens, ask no payment, but record every name in a ledger that never dries. The Daughters are feared because they cut threads unpredictably; a noble may wake to find that a peasant's promise to him has simply \textit{vanished}. The great religions tolerate them because they heal the poor, but the Chain‑Lanterns consider them heretics. Found in Silkstrand, Zakov, and the poorer wards of Ecktoria.

\paragraph{The Cog‑Cults (The Clockwork Monad)}
Underground fraternities of gear‑smiths, logicians, and failed engineers who believe that the universe is a machine that can be perfected. They meet in abandoned watermills, scriptoria, and boiler rooms, reciting the \textit{Canons of Proportion} and performing ``optimisation rites'' on timers, pumps, and looms. The Cog‑Cults are not organised – each cell operates independently – but they share a symbol: a gear with nine teeth (the ninth is hidden). The Everflame condemns them as heretics who worship craft; the Temple of Light ignores them as harmless eccentrics; the Covenant has prosecuted them for ``unauthorised modification of lawful measures.'' Found in Thepyrgos, Aeler, and the industrial wards of Silkstrand.

\paragraph{The Salt Sworn (The Pale Shepherd)}
A syncretic cult of the eastern Dolmis coast, where salt‑panners and fishermen live constantly with the threat of drowning. The Salt Sworn believe that the Pale Shepherd guides not only the dead but also the tide. They build cairns of salt stone at the water's edge and leave no offerings – only the names of the lost, carved into driftwood. Once a year, on the night of the tide of the dead, they chant the names aloud, believing that each name spoken strengthens the threshold between the living and the drowned. The Temple of Light has tried to suppress them three times; each time, the mission ships have been found adrift with no crew and a single name carved into the mast: \textit{Oblivion}. Found in the salt marshes of Zakov, the coast of Theona, and the Aberderrin headlands.

\paragraph{The Bear Cults of Ubral}
In the high glens of Ubral, where the lowland laws of guest‑right and wergild fray against the wilderness, a different faith thrives. The Bear Cults believe that the Savage Heart (whom they call the \textit{Red Maw}) is not a demon but a teacher. Initiates wear cured bear‑skins and carve runes into their ribs with bone needles; they believe that by becoming the predator, they can protect their clans from lowland invaders and the creeping influence of the Everflame. The cult has no holy texts, no clergy, and no temples – only a sacred bear that is fed and honored each autumn. The Lady of Tor has outlawed the Bear Cults three times; each time, the edicts were found torn to shreds outside her hall, pinned with a single claw. Found only in Ubral, rarely encountered by outsiders.

\paragraph{The Ember Prefects (Heretical Everflame)}
Not a separate religion but a persistent schism within the Everflame. The Ember Prefects believe that the flame should not be contained in basilicas – that it must be carried into the streets, into the fields, into the homes of the poor. They meet in arena hypogea and abandoned warehouses, chanting the \textit{Unspoken Ninth Citation} of the Everflame's liturgy (a passage that orthodox priests insist does not exist). Their symbol is a candle that burns from both ends. The Grand Magistrate has outlawed them three times; each time, the edict was found burned the next morning, with a single word left intact: \textit{``Audit.''} The Prefects are hunted by the Inquisitor Prime, but they have sympathizers among the city's poor and among the more radical Hospitaliers. Found primarily in Ecktoria, with small cells in Vhasia and Thepyrgos.

\paragraph{The Veneration of the Hearth (Aelaerem Folk Tradition)}
Not a religion in the institutional sense, but a pervasive set of practices among the halflings of the downs. Every Aelaerem home has a red door, a cup‑mark stone on the threshold, and a pair of watch‑geese. The Hearth is not a god but a presence – the accumulated memory of every kindness shown in that home. Offerings are simple: a crust of bread left on the doorstep, a cup of milk poured into the garden, a red thread tied around the handle of a knife. The Hearth asks for nothing except that guests be welcomed and the ninth cup left unpoured. The Everflame and the Temple of Light have sent missionaries to the downs; they are politely fed, given a bed for the night, and shown the way back to the lowlands in the morning. Their sermons are forgotten by noon. Found only in Aelaerem (Amedell) and among halfling communities in the Dolmis ports.

\paragraph{Player Takeaway}
These traditions are not character classes or mechanical packages; they are sources of story hooks, local color, and potential influence. A character who follows the Grey Pilgrims may gain a way‑cord (treat as a Minor Asset) but owes a lock of hair. A character who joins the Salt Sworn may learn a Hedge Gift but must carve a name into driftwood each full moon. Work with your GM to determine what, if any, mechanical benefit or cost these traditions carry. Most are best represented as Strings, Timers, or narrative permissions, not as full Patron relationships.

\subsection{Cults of the Amaranthine}
\label{subsec:cults}
\index{cults}\index{Cantors}\index{Invokers}\index{Runekeeper}

Where religions build temples, cults gather in shadow. A cult is not a failed religion—it is a \textbf{concentration of purpose}. Where the Everflame concerns itself with all of Ecktoria, a cult of the Inquisitor Prime concerns itself only with \textit{your} heresy. Where the Temple of Light builds hospitals for all, a cult of the Oath of Mercy and Grace builds a single room where promises are forgiven—and \textit{remembered}.

Cults form around a specific Patron, a specific interpretation of that Patron, or a specific living conduit: a \textbf{Cantor} whose voice has drawn the desperate, an \textbf{Invoker} whose bargains have begun to outrun them, or a \textbf{Runekeeper} who has stopped reading the Codex and started writing their own chapters.

\subsubsection*{The Cantor's Congregation}
\index{Cantors!cult leadership}

A Cantor does not choose to lead a cult. A Cantor sings, and the broken listen. The desperate gather. The hungry, the hunted, the hollow—they find the Cantor's song before they find the Cantor. By the time the Cantor realizes they have a congregation, the congregation has already named them \textit{Speaker}, \textit{Shepherd}, or \textit{Storm-Heart}.

A Cantor-led cult is \textbf{charismatic, volatile, and deeply personal}. The Cantor's voice is the cult's scripture; their mood is its liturgy. When the Cantor falters, the cult may fragment, turn inward, or seek a new voice. When the Cantor ascends (or descends), the cult ascends with them—into significance, into terror, or into the annals of inquisitors.

\textbf{GM Seed:} A Cantor of the Pale Shepherd has begun singing funerals for the not-yet-dead. Villagers pay to have their names sung into the Shepherd's flock before they die. The local Temple priest calls it a mercy. The dead call it theft.

\subsubsection*{The Invoker's Covenant}
\index{Invokers!cult leadership}

An Invoker does not seek followers. An Invoker solves problems. But when an Invoker has solved enough problems for enough people, those people begin to \textit{wait}. They wait outside the Invoker's door. They leave offerings. They whisper the Invoker's name alongside the Patron's. A covenant forms—not of worship, but of \textit{dependence}.

An Invoker-led cult is \textbf{pragmatic, secretive, and increasingly brittle}. The Invoker did not ask for loyalty and does not know how to maintain it. Rituals are improvised. Doctrine is whatever worked last time. The cult endures only as long as the Invoker's bargains keep paying out. When they stop, the knives come out.

Invokers themselves are usually \textbf{wary of cults}—cults attract attention, and attention attracts the Inquisitor Prime. But a desperate Invoker may tolerate a covenant, even cultivate it, telling themselves it is just a network, just a favor economy, just a few people who owe them. It never stays \textit{just}.

\textbf{GM Seed:} An Invoker of the Clockwork Monad has accidentally created a cult of gear-makers who believe the Monad speaks through her calculations. They have begun ``optimizing'' the village—replanting fields in perfect grids, scheduling marriages by algorithm. The Invoker cannot stop them without admitting she is not a prophet.

\subsubsection*{The Runekeeper's Schism}
\index{Runekeeper!cult leadership}

A Runekeeper leads a cult only when they have \textit{broken faith}—with their Codex, with their Order, or with the Patron's orthodox interpretation. Runekeeper cults are \textbf{doctrinal, obsessive, and dangerous in proportion to their literacy}. They keep records. They write commentaries. They argue about commas in Rites that should never have been written down at all.

Most Runekeeper cults are small: a master, a few apprentices, a locked scriptorium. But some grow, especially when the Runekeeper claims to have found a \textit{lost Rite}—a fragment of a Patron's power that the orthodox Orders have suppressed. These cults are heresies in the truest sense: not a different faith, but a different \textit{reading} of the same faith. The orthodox hate them most of all.

A Runekeeper's attitude toward cults varies by Patron:
\begin{itemize}
  \item A Runekeeper of the \textbf{Oath of Flame \& Light} sees cults as weeds in the garden—to be pulled, burned, or redirected toward the true flame.
  \item A Runekeeper of \textbf{Ikasha} may tolerate a cult as a useful shadow, but will never stand at its center.
  \item A Runekeeper of \textbf{the Witness} is constitutionally incapable of leading a cult; they would be forced to reveal their own manipulations, and the cult would collapse into truth.
\end{itemize}

\textbf{GM Seed:} A Runekeeper of the Carrion-King has found a Rite that allows followers to ``compost'' their own memories into physical tokens. The cult now trades these tokens—childhoods, first loves, old griefs—on a black market. The Runekeeper insists this is what the King intended. The King has not disagreed.

\subsubsection*{Cults as Factions}

Mechanically, a cult functions as an \textbf{Asset} (see \S\ref{sec:asset-follower-mgmt}) with a severe narrative bias. A cult is never neutral. A cult will:
\begin{itemize}
  \item \textbf{Provide Aid:} Followers can act as Cap 2–5 followers, but only in service of the cult's purpose.
  \item \textbf{Generate Influence:} Leading a cult ticks Influence with the relevant Patron \textbf{+1 per session}—the Patron notices when you gather followers in their name.
  \item \textbf{Attract Attention:} Any cult of 10+ members begins a \textbf{Cult Investigation [6]} timer. When it fills, someone comes looking: an Inquisitor Prime agent, a rival cult, or the Patron themselves.
  \item \textbf{Demand Upkeep:} A cult requires either \textbf{Efficient Upkeep} (XP, representing bribes, shelter, and lies) or \textbf{Intensive Upkeep} (a scene per session where the leader performs a miracle, preaches, or solves a follower's problem).
\end{itemize}

\subsubsection*{Examples of Cults}

The following cults are drawn from the Patron entries throughout this chapter. Each is a template—feel free to adapt, rename, or reskin.

\begin{description}
\item[The Unforgotten Ledger (Varnek Karn)] \hfill \\
  \textit{Funeral notaries who speak with the dead.} They meet in crypts and record every death they witness. Outsiders bring a name and leave with a task.
  \begin{itemize}
    \item \textbf{Leader:} A Death-Speaker (Cantor or Runekeeper).
    \item \textbf{Hazard:} The dead begin to answer—politely, then demandingly.
  \end{itemize}

\item[The Gilded Jest (Mab)] \hfill \\
  \textit{A conspiracy of wits who believe every exchange is a performance.} They meet in theaters and back rooms, trading truths disguised as lies.
  \begin{itemize}
    \item \textbf{Leader:} A Masked Jester (Cantor or Invoker).
    \item \textbf{Hazard:} The cult's contracts become binding fae bargains; breaking one draws Mab's amused attention.
  \end{itemize}

\item[The Iron Chorus (Inquisitor Prime)] \hfill \\
  \textit{Militant witch-hunters who believe purity is survival.} They meet in fortified chapter houses and strike iron bells to seal judgments.
  \begin{itemize}
    \item \textbf{Leader:} A Censor-Prime (Runekeeper or Cantor).
    \item \textbf{Hazard:} The Chorus begins to see heresy everywhere—including in you.
  \end{itemize}

\item[The Compost-Circle (Carrion-King)] \hfill \\
  \textit{A fellowship of gardeners and gravediggers who believe nothing is wasted.} They meet in compost heaps and burial mounds, turning endings into beginnings.
  \begin{itemize}
    \item \textbf{Leader:} A Rot-Warden (Invoker or Hedge-Witch).
    \item \textbf{Hazard:} The Circle's transformations begin to accelerate—composting becomes contagion.
  \end{itemize}
\end{description}

\subsubsection*{Joining or Founding a Cult}

Players may wish to found a cult, join an existing one, or be reluctantly absorbed by one. Use these guidelines:

\begin{itemize}
  \item \textbf{Founding:} Requires a \textbf{Patron's Favor} (a Talent, a critical success on a Rite, or a GM-defined story beat). The new cult begins as a \textbf{Minor Asset} (4 XP) with 1–3 followers.
  \item \textbf{Joining:} Grants access to the cult's \textbf{Asset} benefits (followers, safe houses, information) but ticks Influence \textbf{+1} to the cult's Patron. Leaving a cult is a narrative scene, not a mechanical cost—but the Patron may remember.
  \item \textbf{Reluctant Leadership:} A character who becomes the focus of a cult without seeking it gains the \textbf{Cult Leader} tag. This is a \textbf{Condition}: once per session, the GM may compel the character to act in the cult's interest (preach, protect, perform a Rite) or mark 1 Fatigue from the weight of expectation.
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=black, title=Player Takeaway]
Cults are not for everyone. A cult gives you hands, eyes, and voices beyond your own—but those hands may hold knives, those eyes may watch you sleep, and those voices may begin to speak without your permission. If you choose to lead, remember: the Patron is the real congregation. You are just its favorite hymn.
\end{tcolorbox}

\subsection{Witchcraft}
\label{subsec:witchcraft}

The Amaranthine prays to the Everflame, the Temple, and the Covenant. But when prayers fail, when the child will not be born, when the roads turn against the traveler, the people do not call for a priest. They call for a \textbf{witch}.

A witch is not a heretic—not necessarily. A witch is a \textbf{threshold worker}: one who listens to what has been (Echo), steadies what is (the Veil), and leans into what wishes to be (Flow). Witchcraft is not brute force; it is \textit{fit}—finding where the world already wants to move and giving it a gentle, precise nudge. Hearth craft, mill craft, and road craft are simply slow, patient forms of the same art practiced by ordinary hands. The most powerful magic is often indistinguishable from skilled attention.

\subsubsection{The Two Speeds of Witchcraft}

A witch works at two speeds.

\paragraph{The Long Road}
Ritual, threshold, price, and consequence: curses woven, bargains struck, the world's memory rewritten. It takes \textbf{Significant Time} (hours, days, a moon phase). This is the domain of Full Rituals: unbinding a curse, hallowing a threshold, or unraveling a wound.

\paragraph{The Short Road}
Instinct, residue, and a flicker of attention: a suddenly turned tide, a bell rung at the right moment, a thread plucked. It takes \textbf{Some Time} (a few heartbeats, a single action). This is the domain of \textbf{Quick Workings} and \textbf{Hedge Gifts}—magic so subtle it can be mistaken for luck.

\subsubsection{Core Principles of Witchcraft}

\begin{enumerate}
    \item \textbf{The Law of Named Existence:} To name is to anchor. A True Name fastens an Echo; forgetting frays it. Words spoken under witness reshape how a place remembers.
    \item \textbf{The Concordant Layers:} People, places, and vows exist as chords across Echo / Veil / Flow.
    \begin{itemize}
        \item \textbf{Echo:} Accumulated memory and intention. What has been.
        \item \textbf{Veil:} The current, perceivable state; the boundary between potential and actual.
        \item \textbf{Flow:} The direction of change; the will of elements and intent.
    \end{itemize}
    \item \textbf{Threshold Mechanics:} All workings begin at edges—doorways, wakes, vows, last breaths, first light, crossroads. Magic is boundary-law: who may pass, what must stay, and what should never meet.
    \item \textbf{Element-Will:} Elements are wills, not mere substances. A working \textit{argues} with a chosen will; Influence is the weight incurred when borrowing from a will larger than yours.
    \item \textbf{The Silent Ninth:} The missing step in every scale—absence with teeth. The Ninth is erasure, gaps, omissions, and the power of what is left unsaid. \textbf{Players never declare the Ninth.} The GM may apply a Void tag only as consequence.
\end{enumerate}

\subsubsection{Hedge Gifts}

A \textbf{Hedge Gift} is not a spell. It is the world tilting slightly in the witch's favor because she has paid attention to the right thresholds. Hedge Gifts require no dice roll; they simply work within their limited scope.

Any character with the \textbf{Craft of the Hedge} talent may learn Hedge Gifts. A witch begins with \textbf{two Hedge Gifts} of her choice. Additional Hedge Gifts cost \textbf{2 XP} each.

\begin{description}
    \item[Steady Hand] \hfill \\
    Once per scene, touch a wounded person (including yourself) and whisper a word of comfort. Remove 1 Fatigue. No roll. You may not use this on the same person twice in a scene.

    \item[Knot of Favor] \hfill \\
    Once per scene, tie a small knot in a cord, thread, or cloth. For the rest of the scene, you and allies within Near gain +1 die to a single type of roll (choose when you tie: stealth, persuasion, defense, or navigation). The knot must remain tied; if undone, the benefit ends.

    \item[Warm Hand] \hfill \\
    Once per scene, touch someone and speak a word of comfort or warning. For the next exchange, that person's next physical action either gains +1 die (comfort) or suffers –1 die (warning). No roll. The effect is subtle and cannot be directly perceived as magic.

    \item[Hearth-Sense] \hfill \\
    Once per scene, ask the GM one yes/no question about a threshold, a boundary, or a promise in the current scene. The GM must answer truthfully.

    \item[Salt Line] \hfill \\
    Once per scene, pour a handful of salt in a continuous line (move action). Any spirit, ghost, or cursed creature attempting to cross must test Spirit+Resolve (DV 3) or be stopped for that exchange. The line must be at least Close length; broken if someone steps through or scatters the salt.

    \item[The Unlit Candle] \hfill \\
    Once per session, declare that a light source you can see (lantern, candle, torch) falters or extinguishes. This creates a zone of dim light or darkness in Near range for one exchange.

    \item[Counting the Eighth] \hfill \\
    Once per scene, when you or an ally is about to roll, whisper a number (1–8). If the GM's current Story Beat count is equal to that number modulo 8, the roll gains +2 dice. If not, nothing happens. You cannot use this again until the GM's SB count changes.
\end{description}

\subsubsection{Quick Workings}

A \textbf{Quick Working} is a single-action roll that relies on the witch's knowledge of thresholds, elements, and will. Its effects are limited, subtle, and always carry the risk of consequence.

\paragraph{Procedure}
\begin{enumerate}
    \item \textbf{Name the Threshold} (door, tide line, wound, vow, breath).
    \item \textbf{Choose a Layer} (Echo, Veil, or Flow) and a single Tag (from the magic tags list).
    \item \textbf{Set Position} by urgency (Controlled if you have time, Desperate if threatened).
    \item \textbf{Roll} Attribute + Skill (typically Wits+Lore, Spirit+Insight, or Presence+Sway).
    \item \textbf{Apply Outcome} (see table).
\end{enumerate}

The effect is always minor and scene-limited: a door sticks shut or opens; a candle gutters (brief darkness); a rope knot tightens or loosens; a pursuer's footing slips (1 Fatigue); a whispered name is heard or forgotten.

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Roll Result} & \textbf{Effect} & \textbf{Cost} \\
\hline
Clean Success (Successes $\geq$ DV, SB $=0$) & Effect manifests cleanly. & No cost. \\
\hline
Success with SB (Successes $\geq$ DV, SB $>0$) & Effect works, but GM spends SB for minor complication. & Mark 1 Fatigue or accept a Condition (e.g., Thread-Burned). \\
\hline
Partial ($0 < \text{Successes} < \text{DV}$) & Effect works partially or briefly. & Gain 1 Boon; the threshold remembers you. \\
\hline
Miss ($\text{Successes} = 0$) & Effect fails. & Gain 2 Boons; start a Promise Timer [4] named ``The Threshold Remembers.''\footnote{Promise Timers are an optional mechanic (see the Oath of Mercy and Grace). If not using them, replace with a 4‑segment timer of the same name.} \\
\hline
\end{tabular}
\caption{Quick Working Outcome Matrix}
\end{table}

\subsubsection{Full Rituals}

A Full Ritual requires \textbf{Significant Time}, a proper threshold, and often a witness. Its effects are lasting or permanent: blessings, curses, bindings, or the hallowing of ground.

\paragraph{General Procedure}
\begin{enumerate}
    \item \textbf{Identify the Threshold} (name the edge, learn its disposition).
    \item \textbf{Choose Your Witness} (a person, spirit, or Hollowed who watches).
    \item \textbf{Name Your Will} (what change you intend, using Threshold Tags).
    \item \textbf{Set Your Price} (memory, name, lock of hair, promise, blood).
    \item \textbf{Make the Exchange} (roll, resolve, pay the price on success or partial).
\end{enumerate}

\paragraph{Example Rituals}

\begin{description}
    \item[Unravel the Wound (Heal Harm)] \hfill \\
    \textit{Requires: Significant Time, a threshold (doorway or basin of water), a witness.} Restores 1 level of Harm. \textit{Price:} A lock of hair, a memory of a childhood injury, a tear in a vial. \textit{Roll:} Spirit + Lore vs. DV (3 for Harm 1, 4 for Harm 2, 5+ for Harm 3 requires a coven). \textit{Success:} Harm reduced; witch pays the price. \textit{Partial:} Harm reduced, but witch gains a minor Condition or a Promise Timer [4]. \textit{Miss:} Wound does not heal; witch gains the Condition she tried to remove.

    \item[Hallow Threshold (Permanent Ward)] \hfill \\
    \textit{Requires: Significant Time, a permanent threshold (door, well, bridge, hearthstone), a witness who will maintain the consecration, an offering (bread, salt, coin, spoken promise).} The blessing lasts until the threshold is physically broken or the offering removed. \textit{Roll:} Spirit + Lore vs. DV 4 (home), 5 (village well), or 6 (city gate). \textit{Success:} Threshold hallowed. Spirits, undead, and cursed creatures crossing must test Spirit+Resolve (DV 5) or be turned back. \textit{Partial:} Blessing holds but has a weakness (the witch names one thing that can bypass it). \textit{Miss:} Blessing fails; threshold becomes cursed for one season (all who cross feel unease, –1 die to first action).

    \item[Unbind the Anchor (Permanent Spirit Release)] \hfill \\
    \textit{Requires: Significant Time, the spirit's true name or a strong anchor, a threshold (doorway facing east, crossroads, grave), a witness.} Steps: (1) Discover the anchor (unfinished business, stolen object, broken vow). (2) Name the Terms (propose resolution). (3) Prepare the threshold (salt, chalk, broken cord). (4) Speak the Unbinding (Spirit + Lore vs. DV 4–6). (5) Pay the Price. \textit{Success:} Spirit released. Witch gains 1 Boon. \textit{Partial:} Spirit released but leaves a remnant (echo, cold spot). Witch marks 1 Fatigue and gains a minor Condition. \textit{Miss:} Spirit refuses, becomes hostile. Start a Promise Timer [4] named ``The Spirit's Revenge.''
\end{description}

\subsubsection{Talents of the Hedge}

\begin{description}
    \item[Craft of the Hedge (4 XP)] \hfill \\
    \textit{Requires: Lore 1+ or Survival 1+; a witch mark (tattoo, scar, or birthmark).} You have learned the small magics of thresholds and edges. You gain: two Hedge Gifts; the ability to perform Quick Workings; and once per session, when you fail a Quick Working roll, you may convert the Miss into a Partial Success by marking 1 Fatigue.

    \item[Hedge Gift of the Patron (6 XP)] \hfill \\
    \textit{Requires: Craft of the Hedge, Bond to a Patron.} Once per session, when you would use a Hedge Gift, you may instead invoke your Patron's Imbuement. Gain +1 die to all actions tied to your Patron's domain for the scene, but mark +1 Influence immediately. No action required.
\end{description}

\subsubsection{Covens and Traditions}

As witches practice their craft across the diverse lands of Fate's Edge, distinct traditions have emerged that reflect the unique magical properties of each region. The following are three prominent examples; your GM may introduce others.

\begin{description}
    \item[The Bone Coven (Aelerian Influence)] \hfill \\
    \textit{Focus:} Ancestral wisdom and the patient power of stone that remembers. \textit{Signature Rite:} Stone Memory Communion—by pressing hands to ancient stone, the witch witnesses events that occurred near it, going back centuries. \textit{Patron Elements:} Earth and Fate.

    \item[The Mistwalkers (Mistlands Influence)] \hfill \\
    \textit{Focus:} Navigation between worlds and manipulation of thresholds. \textit{Signature Rite:} Wayfinding Through the Veil—by burning specific herbs, the witch creates temporary passages through the Ways Between. \textit{Patron Elements:} Water and Obishaal (Death/Dreams).

    \item[The Stormweavers (Linn/Ykrul Influence)] \hfill \\
    \textit{Focus:} Weather magic and storm riding. \textit{Signature Rite:} Storm Dancing—through ritual movement, the witch calls forth or calms storms. The most experienced can ride the winds themselves. \textit{Patron Element:} Air, with strong ties to Raéyn or Xhak'Thul.
\end{description}

\subsubsection{Threshold Magic and Promise Timers}

Many workings create obligations. Represent these with a \textbf{Promise Timer } (typically 4 segments):

\begin{itemize}
    \item Mark the timer when the promise is ignored, deferred, or disrespected.
    \item When filled, the price comes due—socially, spiritually, or narratively.
    \item Examples: sharing grief, tending a grave, protecting a hearth, reconciling rivals.
\end{itemize}

\paragraph{Household Disposition}
Communities and dwellings possess a temperament. When entering a new settlement or threshold site, the GM may choose or roll for its disposition:

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Disposition} & \textbf{Effect} \\
\hline
Hungry & Demands offerings; first working costs +1 Fatigue. \\
Grieving & Magic requires memory or loss narrative. \\
Fearful & First rite is Desperate Position. \\
Watching & +1 Effect if the house is fed or soothed. \\
Young & Chaotic outcomes on partials. \\
Old & Weak boons but harsh backlash. \\
\hline
\end{tabular}
\end{center}

Household temperament can change through care, ritual, or story.

\subsubsection{The Daughters of the Unbroken Chord (The Cord)}

Among the most feared and misunderstood covens are the \textbf{Daughters of the Unbroken Chord}—also called the Grey Watchers, Pale Midwens, or simply the Chain. They appear as midwives, hospice nurses, and deathbed witnesses, offering the mercy of warm hands and the comfort of final words. They seem harmless. Mostly.

The Daughters are bound to the \textbf{High Mother}, a threshold entity sustained by accumulated promise and subtle sacrifice. She may be a lich. She may be the current Vessel: a mortal woman elected by the coven, granted godlike power through centuries of accumulated influence. The Daughters smile when asked which is true. The Cord does not distinguish between memory and flesh.

\paragraph{Becoming a Daughter}
A character may join the Daughters of the Cord through the following requirements:
\begin{itemize}
    \item \textbf{Attributes:} Spirit 3+, Body 2+
    \item \textbf{Skills:} Lore (Thresholds) 1+, Subterfuge 1+
    \item \textbf{Talents:} Disciplined Body (Monk) [3 XP], Grey Thread [4 XP] (see below)
\end{itemize}

\begin{description}
    \item[Grey Thread (4 XP)] \hfill \\
    \textit{Requires: Disciplined Body, Spirit 3+.} You are bound to the High Mother's Cord. Your flesh accepts tattooed Rites; your blood activates them. You may perform Rites of the High Mother without a Codex or Thiasos. Instead, you receive a \textbf{Hollowed}—a heartless familiar with eyes like spilled milk. It witnesses your magic and counts as the threefold presence required by Hag Law. The Hollowed never speaks.

    \item[Cord of Flesh (8 XP)] \hfill \\
    \textit{Requires: Grey Thread, Body 3+.} Your body stores memories and promises. When you convert Harm $\rightarrow$ Fatigue (Monk ability), you may store up to Body rating points of Fatigue as \textbf{Chained Knots}, appearing uninjured (as fresh scars or weeping tattoos). You have until sunset after the next to perform a sacrifice or take 2 Harm.

    \item[The Open Palm, Closed Cord (8 XP)] \hfill \\
    \textit{Requires: Grey Thread, Melee 2+.} Your unarmed strikes or subtle touch carry \textbf{Widow's Measure}. Instead of dealing Harm, you may mark a Promise Timer [4] on the target—they owe the Cord a truth or grief. If they already owe the coven, you may activate that weight: they suffer Position –1 for the scene, or you gain Effect +1 against them.
\end{description}

\paragraph{The Subtle Sacrifice}
The Daughters do not slaughter goats on altars. They require \textbf{domestic sacrifices}—acts of subtlety so small they are dismissed as accidents or cruelty, yet sufficient to feed the Cord.

\begin{itemize}
    \item \textbf{The Missing Child:} Taken to the House Without Doors. The child is hollowed out into a Heartless familiar. The parents forget they existed (Un-Naming).
    \item \textbf{The Unspoken Grief:} At a deathbed, the Daughter convinces the dying to forget their greatest love. The memory is trapped in a Name-Bead. The dying smile peacefully, having forgotten why they were sad.
\end{itemize}

Performing a sacrifice clears 1 segment of Influence with the High Mother or reduces a Rite's cost by 1 (minimum 1). Each requires \textbf{Intricate Description} and leaves physical evidence.

\begin{tcolorbox}[colback=gray!5!white, colframe=black, title=Player Takeaway]
Witchcraft is not a separate magic system—it is a \textbf{way of working} that any character with the right talents can learn. A Runekeeper can learn Hedge Gifts. An Invoker can perform a Quick Working. A character with no Patron at all can become a capable hedge-witch, tending thresholds and binding small promises. The power of witchcraft is not in its flash, but in its patience. The world remembers. And the witch knows where to touch it.
\end{tcolorbox}