\section{Living Thiasos — Creatures of Flesh and Bone}

A living Thiasos breathes, eats, sleeps, and bleeds. It is a creature of flesh and bone, loyal to its Runekeeper but possessed of its own instincts, fears, and preferences. Unlike a construct or spirit, a living Thiasos cannot be repaired—it must be healed, rested, and cared for. But it offers something that no machine or ghost can: the warmth of a living bond.

\subsection{Warhorse}

\noindent\textbf{Patrons:} Oath of Flame and Light, Aveh, The Traveler, Oath of Mercy and Grace

\noindent\textbf{Appearance:} A horse of exceptional breeding—grey for the Oath of Mercy, black for Aveh, white for the Oath of Flame. Its eyes are calm. Its hooves are steel-shod. It does not startle at fire or the clash of steel.

\noindent\textbf{Cap:} 3 (advances to 5)

\noindent\textbf{Harm:} 5 (advances to 8)

\noindent\textbf{Traits:}

\begin{itemize}
\item \textit{Swift:} +1 die to chase or escape rolls when mounted.
\item \textit{Trample:} When the Runekeeper charges, the warhorse may also attack (Harm 2 to one target in Close range).
\item \textit{Unspookable:} The warhorse does not need to test Resolve against fear caused by fire, combat, or loud noises.
\item \textit{Loyal (Advanced):} The warhorse will not flee while the Runekeeper is mounted, even if wounded.
\end{itemize}

\noindent\textbf{Advanced Traits (Tier II+):}
\begin{itemize}
\item \textit{Phantom Hooves (Tier II):} The warhorse's hooves leave no tracks for one hour per day.
\item \textit{Spirit Steed (Tier III):} The warhorse can gallop across water as if it were solid ground for up to one minute.
\end{itemize}

\subsection{Spirit-Hound}

\noindent\textbf{Patrons:} Mykkiel, The Sealed Gate, Varnek Karn, Inquisitor Prime

\noindent\textbf{Appearance:} A hound of grey or black fur, eyes like polished slate or burning coals. It moves silently. It does not bark unless commanded.

\noindent\textbf{Cap:} 2 (advances to 4)

\noindent\textbf{Harm:} 4 (advances to 7)

\noindent\textbf{Traits:}

\begin{itemize}
\item \textit{Tracker:} +2 dice to track a target by scent.
\item \textit{Spirit Sense:} Can detect invisible or incorporeal creatures within Near range.
\item \textit{Bound to Oath:} If the Runekeeper breaks a sworn oath, the spirit-hound whines and refuses to act for one scene.
\item \textit{Silent Step:} The hound makes no sound when moving, even across dry leaves or loose stone.
\end{itemize}

\noindent\textbf{Advanced Traits (Tier II+):}
\begin{itemize}
\item \textit{Oathbreaker's Bane (Tier II):} The hound gains +1 die to track any creature that has broken an oath witnessed by the Runekeeper.
\item \textit{Phantom Bite (Tier III):} The hound's bite harms incorporeal creatures as if they were flesh.
\end{itemize}

\subsection{Wolf-Ancestor}

\noindent\textbf{Patrons:} Grimmir, The Savage Heart, Ykrul traditions

\noindent\textbf{Appearance:} A grey wolf the size of a pony, with eyes that reflect scenes from the Runekeeper's childhood. Its fur is thick; its teeth are white.

\noindent\textbf{Cap:} 3 (advances to 5)

\noindent\textbf{Harm:} 4 (advances to 7)

\noindent\textbf{Traits:}

\begin{itemize}
\item \textit{Pack Hunter:} When the Runekeeper attacks a target, the wolf-ancestor gains +1 die to attack the same target.
\item \textit{Howl:} Once per scene, the wolf-ancestor howls. Allies gain +1 die to resist fear for the scene.
\item \textit{Ancestral Memory:} The wolf-ancestor recalls the names of the Runekeeper's ancestors (GM may provide hints or lore).
\item \textit{Blood Trail:} The wolf can track any creature that has bled within the last hour.
\end{itemize}

\noindent\textbf{Advanced Traits (Tier II+):}
\begin{itemize}
\item \textit{Ancestral Fury (Tier II):} When the Runekeeper is reduced to 0 Harm, the wolf gains +2 dice to all actions for the rest of the scene.
\item \textit{Spirit Walk (Tier III):} The wolf can become incorporeal for one exchange, passing through walls and obstacles.
\end{itemize}

\subsection{Fungal Familiar}

\noindent\textbf{Patrons:} The Carrion-King, Grimmir (rot aspect)

\noindent\textbf{Appearance:} A cluster of pale mushrooms growing from a crack in the Runekeeper's armor, or a small fungal mass that rides on the shoulder. It pulses faintly. It smells of damp earth.

\noindent\textbf{Cap:} 2 (advances to 3)

\noindent\textbf{Harm:} 3 (advances to 5)

\noindent\textbf{Traits:}

\begin{itemize}
\item \textit{Spore Cloud:} Once per scene, the familiar releases a cloud of spores. Enemies in Close range test Endurance (DV 3) or suffer -1 die to physical actions for one exchange.
\item \textit{Decay Sense:} The familiar can detect rot, decay, or corruption within Near range.
\item \textit{Regenerative:} The familiar regains 1 Harm per day (faster than natural healing).
\item \textit{Spore Messenger:} The familiar can release a single spore that drifts to a target within Far range, carrying a one-sentence message.
\end{itemize}

\noindent\textbf{Advanced Traits (Tier II+):}
\begin{itemize}
\item \textit{Corpse Bloom (Tier II):} The familiar can accelerate decay in a corpse, reducing it to compost in minutes.
\item \textit{Spore Possession (Tier III):} The familiar can release a cloud that briefly animates a corpse (Cap 1, one action).
\end{itemize}

\subsection{Great Cat Spirit}

\noindent\textbf{Patrons:} The Savage Heart, Inaea (shadow aspect)

\noindent\textbf{Appearance:} A translucent leopard, lion, or tiger, visible only in moonlight or shadow. Its eyes are amber. Its paws make no sound.

\noindent\textbf{Cap:} 3 (advances to 4)

\noindent\textbf{Harm:} 4 (advances to 6)

\noindent\textbf{Traits:}

\begin{itemize}
\item \textit{Stalker:} The cat gains +2 dice to Stealth in natural or urban terrain.
\item \textit{First Strike:} When the cat attacks from hiding, its target suffers -1 die to defend.
\item \textit{Shadow Step:} The cat can move through shadows as if they were open space, ignoring difficult terrain caused by darkness.
\item \textit{Silent Kill:} The cat can make a killing blow without sound—useful for assassinations.
\end{itemize}

\noindent\textbf{Advanced Traits (Tier II+):}
\begin{itemize}
\item \textit{Phantom Pounce (Tier II):} The cat can attack from hiding at Far range, closing the distance in the same action.
\item \textit{Terrifying Roar (Tier III):} Once per session, the cat roars. All enemies within Near range test Resolve (DV 4) or become Shaken.
\end{itemize}

\subsection{Hollowed Child (or Heartless Bird)}

\noindent\textbf{Patrons:} The High Mother

\noindent\textbf{Appearance:} A child of perhaps seven years, pale and still, with milk-white eyes sewn open with silver wire. It does not blink. It does not speak. It stands in corners and watches. Alternatively, a small grey bird with the same milk-white eyes, perched on the Runekeeper's shoulder.

\noindent\textbf{Cap:} 2 (does not advance)

\noindent\textbf{Harm:} Cannot be harmed by mundane means

\noindent\textbf{Traits:}

\begin{itemize}
\item \textit{Silent Witness:} The Hollowed counts as a witness for all Rites. It never blinks.
\item \textit{Memory Keeper:} The Hollowed records every debt it witnesses. The Runekeeper may consult it to recall a forgotten promise or obligation.
\item \textit{Unsettling Presence:} Animals and children are disturbed by the Hollowed. Social rolls involving them suffer -1 die if the Hollowed is visible.
\item \textit{Blood Tears:} When a debt comes due, the Hollowed weeps blood. The blood can be used as a component in Rites of the High Mother.
\end{itemize}

\noindent\textbf{Advanced Traits (None):} The Hollowed does not advance. It is eternal—and unchanging.

\subsection{Selkie-Spirit}

\noindent\textbf{Patrons:} Raéyn, Linn traditions

\noindent\textbf{Appearance:} A seal that can shed its skin to become a woman with hair like kelp and eyes like storm clouds. The skin is grey and slick; the human form is pale and wet.

\noindent\textbf{Cap:} 3 (advances to 4)

\noindent\textbf{Harm:} 4 (advances to 6)

\noindent\textbf{Traits:}

\begin{itemize}
\item \textit{Amphibious:} The selkie can breathe underwater and move at full speed on land or sea.
\item \textit{Tide Reader:} The selkie knows the currents within Near range—useful for navigation and escape.
\item \textit{Storm Sense:} The selkie can predict the weather for the next hour with perfect accuracy.
\item \textit{Drowning Grasp:} The selkie can pull a target underwater (if available) and hold them there, dealing 1 Fatigue per exchange.
\end{itemize}

\noindent\textbf{Advanced Traits (Tier II+):}
\begin{itemize}
\item \textit{Shape Change (Tier II):} The selkie can shift between seal and human form as a free action.
\item \textit{Call the Tide (Tier III):} Once per session, the selkie can summon a wave that pushes all targets in a Near zone one band away.
\end{itemize}

\subsection{General Care and Maintenance}

Unlike constructs or spirits, living Thiasos require:

\begin{itemize}
\item \textbf{Food and Water:} Daily, or they suffer -1 die to all actions.
\item \textbf{Rest:} At least four hours per day, or they become Winded (-1 die to all actions).
\item \textbf{Healing:} Rest, Medicine, or magical healing. A living Thiasos cannot be repaired—only healed.
\end{itemize}

Neglecting a living Thiasos may cause it to become Wary (hesitates to assist) or, in extreme cases, to leave the Runekeeper's service entirely.

\subsection{Living Thiasos and Obligation}

When the Runekeeper marks Obligation, a living Thiasos may also bear a visible mark:

\begin{itemize}
\item A warhorse develops a limp.
\item A spirit-hound's glow dims.
\item A wolf-ancestor loses a patch of fur.
\item A fungal familiar's mushrooms wilt.
\item A great cat spirit becomes translucent.
\item A selkie's skin grows dry.
\end{itemize}

The mark fades when the Obligation is cleared. The Thiasos does not forget.

\subsection{Living Thiasos and Death}

If a living Thiasos dies, the Runekeeper marks 2 Obligation immediately and gains the Condition \textit{Bereaved} (-1 die to all rolls involving trust or loyalty) until they acquire a new Thiasos. A funeral is optional—but recommended. The dead remember.

The thread held. The water carried. That is not a Rite. That is grace.
