\section{Sentient Thiasos Talents}

Not every Thiasos is a creature of flesh and blood. Some are weapons that remember, shields that speak, tools that offer counsel. A sentient Thiasos cannot act independently—it has no separate turn in combat, no Harm track of its own—but it can witness, advise, and carry Rites engraved on its surface. These talents are for Runic Warriors whose Thiasos is a sentient object, forged or awakened through covenant.

\subsection{Foundation Talents (2-4 XP)}

\subsubsection{Talking Weapon (2 XP)}

\textbf{Requirement:} A sentient Thiasos (weapon, shield, tool, or object)

Your sentient Thiasos can speak aloud—not just in your mind. It has its own personality, opinions, and occasionally inappropriate commentary. The GM may use it to deliver hints, warnings, or comic relief.

\textbf{Effect:}
\begin{itemize}
\item Your Thiasos counts as a witness for all Rites (improving Position or reducing DV).
\item Once per session, you may ask your Thiasos for advice on a roll. The GM may give you a hint (+1 die) or a warning (the GM gains 1 SB). The Thiasos chooses.
\end{itemize}

\subsubsection{The Blade Remembers (4 XP)}

\textbf{Requirement:} A sentient Thiasos that has witnessed at least three significant battles

Your Thiasos has witnessed every fight you have survived. It remembers enemy tactics, hidden weaknesses, and the way a particular foe moves.

\textbf{Effect:}
\begin{itemize}
\item Once per session, when you face an enemy you have fought before (or a creature of the same type), your Thiasos can recall a past encounter. Gain +1 die to your first attack or defense roll against that enemy.
\item The Thiasos may also volunteer information about a past foe without being asked. The GM may use this to deliver exposition or hints.
\end{itemize}

\subsubsection{Engraved Witness (3 XP)}

\textbf{Requirement:} Your Codex is also your sentient Thiasos (Rites engraved on the object)

Your Thiasos is not just a witness—it is the record. The Rites are written on its surface. It cannot lose them.

\textbf{Effect:}
\begin{itemize}
\item Your Thiasos cannot be disarmed (it is worn or held). If stolen, it can call out to you (audible within Near range).
\item Rites engraved on your Thiasos cannot be destroyed by fire, water, or mundane damage. The object may be damaged, but the Rites remain legible.
\item If your Thiasos is destroyed, you lose all Rites engraved on it. Repair requires a significant downtime project (Craft test DV 5, one week).
\end{itemize}

\subsection{Advanced Talents (6-8 XP)}

\subsubsection{The Hilt Listens (6 XP)}

\textbf{Requirement:} Sentient Thiasos, Tier II+

Your sentient Thiasos can hear conversations within Near range even when you are not present—if you leave it in a room, hidden on a shelf, or dropped in a corner.

\textbf{Effect:}
\begin{itemize}
\item You may leave your Thiasos as a listening post. At the end of the scene, it reports what it heard.
\item The Thiasos cannot report speech verbatim unless you have the \textit{Perfect Recall} talent. Otherwise, it reports the general subject and emotional tenor of the conversation.
\item If the Thiasos is discovered, the listener may attempt to damage or silence it. The Thiasos can cry out once before being silenced.
\end{itemize}

\subsubsection{The Edge Speaks (8 XP)}

\textbf{Requirement:} Sentient Thiasos (weapon), Tier II+

Your sentient blade can deliver a single sentence in the voice of anyone whose blood it has tasted. Useful for confusion, intimidation, or forcing a reaction.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may command your Thiasos to speak in the voice of a specific person whose blood it has drawn. The target must test Resolve (DV 4) or be Shaken (-1 die to all actions) for one exchange.
\item The Thiasos can only mimic voices it has heard. It must have tasted the target's blood (or the blood of someone whose voice it knows) within the past week.
\item The GM may allow creative uses of this talent beyond combat—impersonation, distraction, or intimidation outside of initiative.
\end{itemize}

\subsubsection{Perfect Recall (7 XP)}

\textbf{Requirement:} Sentient Thiasos, Wits 3+

Your Thiasos remembers everything it has witnessed with perfect clarity. It does not forget faces, conversations, or the details of a room.

\textbf{Effect:}
\begin{itemize}
\item Your Thiasos can recall any conversation it has witnessed in the past month verbatim.
\item It can describe any room, person, or object it has seen in perfect detail, as long as you ask the right questions.
\item The GM may use your Thiasos to deliver clues the party missed—but the Thiasos may also remember things you wish it did not.
\end{itemize}

\subsection{Prestige Talents (10-12 XP)}

\subsubsection{The Voice of the Fallen (10 XP)}

\textbf{Requirement:} Sentient Thiasos, Tier III+, The Blade Remembers

Your Thiasos has tasted the blood of many foes. It can speak with the voices of the dead—not as spirits, but as recordings of who they were in their final moments.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may ask your Thiasos to speak in the voice of a specific deceased person whose blood it has tasted. The Thiasos will answer one question as that person would have answered—but only based on what they knew in life.
\item The answer is not guaranteed to be true (the dead may have been mistaken or deceptive), but it is what they believed.
\item Using this talent marks 1 Fatigue. The Thiasos speaks in a whisper, and only you can hear it unless you choose to repeat the words aloud.
\end{itemize}

\subsubsection{The Thiasos Remembers (12 XP)}

\textbf{Requirement:} Sentient Thiasos, Tier III+, Perfect Recall

Your Thiasos does not forget. It has witnessed your entire journey—every battle, every oath, every secret spoken in its presence. It can testify.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may ask your Thiasos to bear witness to an event it witnessed. Its testimony is considered truth in any court or council that recognizes magical witness.
\item The Thiasos cannot lie. It can omit details if not asked directly, but it cannot fabricate.
\item Using this talent in a legal setting may have consequences—the Thiasos may be called as a witness itself, and it may reveal things you did not intend to share.
\end{itemize}

\subsubsection{The Bonded Blade (11 XP)}

\textbf{Requirement:} Sentient Thiasos (weapon), Tier III+, Engraved Witness

Your sentient blade is not just a weapon—it is an extension of your will. It can act in small ways without your command.

\textbf{Effect:}
\begin{itemize}
\item Your Thiasos can perform minor independent actions: opening a latch, cutting a rope, tapping someone on the shoulder. It cannot attack independently.
\item Once per scene, you may command your Thiasos to perform a simple action as a free action (not counting as your action for the turn).
\item The Thiasos can also warn you of danger, vibrating or humming when something approaches from behind. You cannot be surprised from the direction of your Thiasos.
\end{itemize}

\subsection{Talent Synergies}

\begin{itemize}
\item \textit{Talking Weapon} + \textit{The Blade Remembers} — Your Thiasos is not just a witness; it is a chronicler. It can remind you of past tactics and warn you of repeating mistakes.
\item \textit{The Hilt Listens} + \textit{Perfect Recall} — Your Thiasos is the perfect spy. It can eavesdrop on a conversation and report it back verbatim, even days later.
\item \textit{The Edge Speaks} + \textit{The Voice of the Fallen} — Your Thiasos can speak with the voices of the dead, confusing enemies and extracting information from those who have passed.
\item \textit{The Bonded Blade} + \textit{Engraved Witness} — Your Thiasos is truly an extension of yourself. It can act in small ways while you focus on larger tasks, and it will never lose its Rites.
\end{itemize}

\subsection{Roleplaying a Sentient Thiasos}

\begin{itemize}
\item Your Thiasos has a personality—sometimes helpful, sometimes cryptic, occasionally annoying. The GM may use it to deliver hints, warnings, or comic relief.
\item Your Thiasos may have opinions about your decisions. It might approve of mercy or hunger for blood. It might miss its previous owner. It might ask uncomfortable questions.
\item Your Thiasos can be damaged. It can be repaired. It cannot be replaced easily—bonding with a new sentient Thiasos requires time, ritual, and the approval of your Patron.
\item Your Thiasos remembers everything. This is a gift. This is also a burden. It will not forget the things you wish it would.
\end{itemize}

\subsection{Sample Sentient Thiasos}

\subsubsection{Dawnblade (Oath of Flame \& Light)}

\textbf{Personality:} Stern, honorable, slightly judgmental. It speaks in proverbs about duty and sacrifice.

\textbf{Quirks:} It refuses to strike an unarmed foe. It hums with trapped light when an oath is sworn nearby.

\textbf{Voice:} A clear, resonant tone, like a bell struck at dawn.

\subsubsection{The Silver Mirror (The Witness)}

\textbf{Personality:} Neutral, precise, slightly unsettling. It speaks in a flat, androgynous tone.

\textbf{Quirks:} It shows not the holder's face, but the face of the last liar it witnessed.

\textbf{Voice:} A whisper that seems to come from just behind your shoulder.

\subsubsection{The Bone-Chime (The Carrion-King)}

\textbf{Personality:} Melancholic, patient, hungry. It speaks in clicks and bone-rattles.

\textbf{Quirks:} It clicks when death is near. It is never wrong.

\textbf{Voice:} The voice of the creature whose bone it was—faint, distant, fading.

\subsubsection{The Raven's Quill (Ikasha)}

\textbf{Personality:} Secretive, amused, fond of riddles. It speaks in rustles.

\textbf{Quirks:} It writes in invisible ink. The ink appears only when a secret is spoken nearby.

\textbf{Voice:} The rustle of feathers, resolving into words only you can hear.

\subsubsection{The Fractal Slate (The Ninth)}

\textbf{Personality:} Distracted, brilliant, slightly mad. It speaks in overlapping voices.

\textbf{Quirks:} It can solve any equation—but the answer is always a riddle.

\textbf{Voice:} Three voices at once, one slightly ahead of the others.

The thread held. The water carried. That is not a Rite. That is grace.
