\section{The Ninth — Beyond Comprehension}

\subsection{Domain and Epithets}

Beyond the eighth figure of Sacred Geometry lies a rim no compass will hold. The Ninth is not absence but surplus: knowledge poured past the lip of any vessel, truth that overgrows the frame that seeks to bind it. Those who bend the knee at its lectern become conduits for impossible reckonings—glimpsing patterns that cannot be kept whole inside a single skull. Each revelation purchases power at the price of living comfortably within the world's common measures.

The Ninth does not explain, define, or warn. It overflows. To invoke it is to step past the eighth threshold—the one you did not know was there—and find that on the other side, the cup forgets its shape, the bell rings a note that was never cast, and the answer to your question arrives as a question you had not yet learned to ask.

\subsection{Cultural Origins}

The Ninth is acknowledged wherever knowledge exceeds containment:

\begin{itemize}
\item In \textbf{Thepyrgos}, the Ninth is the Lost Syllable. Archivists believe it is a word erased from every dictionary—but still whispered in dreams.
\item In \textbf{Aelinnel}, the Ninth is the Missing Number. Geometers speak of a quantity that cannot be written, only approximated—and never safely.
\item In \textbf{Aeler}, the Ninth is the Deep Count. Vent-priors warn that some ledgers should never be balanced; the ninth column is for the mountain's own debts.
\item In \textbf{Lethai} courts, the Ninth is the Unwritten Clause—the term that exists only in the silence between spoken words.
\end{itemize}

\subsection{Warrior Archetype: The Fractal Blade}

A Runic Warrior of the Ninth is not a keeper of secrets or a collector of debts. They are a \textit{Fractal Blade}—a warrior who fights with impossible angles, who sees the pattern beneath the chaos, whose strikes land where no strike should be possible. Their Codex is a broken mirror or a fractal slate. Their Thiasos is a pattern-spirit that whispers probabilities, a fractured echo that shows alternate outcomes, or a sentient blade that exists in two places at once.

\subsection{Codex Form}

The Fractal Blade's Codex is a testament to broken geometry:

\begin{itemize}
\item \textbf{Broken Mirror:} A hand mirror shattered into three pieces, bound with silver wire. Each shard shows a different angle of the same truth.
\item \textbf{Fractal Slate:} A slate tablet covered in spiraling, unfinished equations. The equations change when no one is looking.
\item \textbf{Unfinished Compass:} A brass compass whose needle spins without stopping. It points not to north, but to the nearest paradox.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Pattern-Spirit:} A translucent form of shifting geometry—triangles that become squares, spirals that reverse. It can predict the outcome of a single roll (GM must answer "success" or "failure" before the roll).
\item \textbf{Fractured Echo:} A shimmering duplicate of the Runekeeper that exists half a second in the future. It can show an alternate outcome of an action—but cannot act.
\item \textbf{The Fractal Slate (Sentient):} A slate that speaks in overlapping voices. It can solve any equation—but the answer is always a riddle.
\end{itemize}

\subsection{Core Mechanic: The Glimpse}

Once per scene, the Fractal Blade may glimpse a shard of Ninth truth. Choose one: gain +1 die on your next roll, ask the GM one yes/no question about a hidden pattern, or treat a failed roll as a Partial Success. After glimpsing, mark 1 Fatigue—the knowledge strains the skull.

When the Fractal Blade rolls a critical success, they may ask the GM one question about the underlying pattern of the scene. The GM must answer truthfully, though the answer may be cryptic.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Ninth Figure (4 XP)}

\textbf{Tags:} [KNOWLEDGE], [THRESHOLD]

\textbf{Threshold:} A chalk figure with an "impossible" joint, or a reckoning worked toward infinity on slate.

Glimpse past the ordinary frame of measure. Gain +1 die to Lore or Investigation when wrestling complex systems or many-layered problems.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Flooded Scriptorium (5 XP)}

\textbf{Tags:} [KNOWLEDGE], [SPEED], [OVERFLOW]

\textbf{Threshold:} An overburdened ledger-hall, stacked codices, a bell that tolls the copying-hour without cease.

Read faster than hands can turn pages. Ask three questions upon any topic that leaves a written wake; gain +2 dice to rolls of analysis this scene.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Hypercognitive Lattice (8 XP)}

\textbf{Tags:} [BOOST], [THRESHOLD], [OVERFLOW]

\textbf{Threshold:} A brass-thought lattice, an orrery of inked threads, or a draught that sharpens memory.

Widen the mind's mouth. For this scene, treat your Wits as +2 higher for information-processing rolls; you may track several intricate streams at once.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Breaking Glyph (7 XP)}

\textbf{Tags:} [KNOWLEDGE], [TRUTH], [THRESHOLD]

\textbf{Threshold:} A vigil in dim light, a water-clock, and a problem sworn not to be left half-seen.

Take one pattern whole. Choose a complex system, theory, or design and gain perfect internal grasp of it; receive +3 dice to related knowledge rolls for the session.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Ninth Revelation (14 XP)}

\textbf{Tags:} [TRUTH], [THRESHOLD], [TRANSFORM], [FATE]

\textbf{Threshold:} A fast from easy speech, a circle of unbroken figures, and vows to see past comfort.

Glimpse a shard of Ninth truth. Take one major insight that overleaps common frames—perfect command of a knotty system, forecast of chaotic motion, or comprehension of an "impossible" relation.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Glimpse Beyond (4 XP)}

\textbf{Requirement:} The Ninth

Once per session, look at a complex situation (a crime scene, a puzzle lock, a battle formation) and instantly intuit the single most important hidden detail. The GM gives you one piece of information that would otherwise require a successful Investigation roll. Mark 1 Fatigue after using this talent.

\subsubsection{Pattern Recognition (6 XP)}

\textbf{Requirement:} The Ninth

Once per scene, see connections others miss within tangled accounts and signs. You gain +1 die to any roll involving pattern recognition, system analysis, or prediction.

\subsubsection{Fractal Response (8 XP)}

\textbf{Requirement:} The Ninth, Wits 3+

When an enemy attacks you, you may glimpse the outcome a half-second in advance. Your defense roll gains +1 die. You may use this once per scene.

\subsection{Roleplaying a Fractal Blade}

\begin{itemize}
\item You do not see the world as others see it. You see patterns beneath patterns. The geometry is always there—most people simply refuse to look.
\item You may be called to solve the unsolvable, to find the hidden flaw, to predict the unpredictable. The Ninth does not give answers. It gives angles.
\item You may struggle with the cost of knowing. Every glimpse leaves a mark. Every insight steals a piece of comfortable ignorance. You cannot unsee what you have seen.
\item You may be asked to stop—to close the book, to put down the mirror, to forget the equation. The Ninth does not offer forgetting. It offers more.
\end{itemize}

\subsection{Sample NPC: Geometer Tolin of the Unfinished Compass}

\textbf{Archetype:} Fractal Blade

\textbf{Patron:} The Ninth

\textbf{Codex:} A broken mirror bound with silver wire

\textbf{Thiasos:} A pattern-spirit that whispers probabilities

\textbf{Personality:} Distracted, brilliant, fraying. He calculated the ninth digit of pi once. He has not slept since.

\textbf{Conflict:} He has seen the pattern underlying the war—the geometry of loss, the equation of victory. If he speaks it aloud, the war could end in a day. The cost is that the pattern will see him too. The Ninth is watching.

\subsection{GM Guidance: Intrusions for The Ninth}

\begin{itemize}
\item The Fractal Blade sees patterns in everything—the way dust falls, the arrangement of stones, the spacing of footsteps. It is useful, but exhausting. Mark 1 Fatigue.
\item A fact the Fractal Blade learned from the Ninth is true, but should not be. Others find it disturbing when they speak it aloud.
\item The Fractal Blade begins to forget mundane things (names, faces, what they ate) while remembering impossible geometries with perfect clarity.
\item The Fractal Blade dreams of an equation that has no solution. When they wake, they cannot speak for one hour—the numbers are still on their tongue.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
