\section{Lunara — The Silver Quiet}

\subsection{Domain and Epithets}

Lunara walks where torchlight falters: crossroads, tidal edges, shadowed sanctuaries. She is the whisper beneath silver skies, a torch-bearing guide to mysteries that lie between waking and dream. The ancients know her as triple-formed: maiden of new light, mother of full radiance, crone of waning shadow. Her worshippers—night-scholars, dream-walkers, and hedge-priests—seek her not for power over others, but for passage, reflection, and revelation.

Her gifts are gentle yet perilous: a mirror that reveals as much of the seeker as the sought, a key that opens ways better left closed, a light that makes the unseen plain. Her sigil is the silver spiral, etched in water, mirror, or stone.

\subsection{Cultural Origins}

Lunara is honored wherever moonlight touches threshold and secret:

\begin{itemize}
\item In \textbf{Aelinnel}, she is the Silver Path, invoked by those who walk the Ways Between. Her shrines are moonlit groves where no iron may enter.
\item In \textbf{Lethai} courts, she is the Mirror-Watcher, whose priests read futures in still water. Her symbol is a crescent moon reflected in a pool.
\item In \textbf{Theona}, she is the Quiet Sister, invoked by those who must cross the Green Host's territory on the night of the full moon.
\item In \textbf{Vhasia}, she is the Lantern-Bearer's Shadow, the light that guides lost souls—and the dark that hides them from those who hunt.
\end{itemize}

\subsection{Warrior Archetype: The Moon Knight}

A Runic Warrior of Lunara is not a weaver of webs or a collector of debts. They are a \textit{Moon Knight}—a warrior who fights under moonlight, whose strength waxes and wanes with the silver sphere, who hunts the creatures that fear the light and protects those who walk in darkness. Their Codex is a silver disc or a moon-bell. Their Thiasos is a shadow-cat that moves without sound, a lunar spirit that reveals hidden thresholds, or a sentient blade that glows under moonlight.

\subsection{Codex Form}

The Moon Knight's Codex is a testament to silver and shadow:

\begin{itemize}
\item \textbf{Silver Disc:} A polished disc of silver, warm to the touch. It glows faintly under moonlight. In darkness, it is cold—a warning.
\item \textbf{Moondrop Lens:} A lens of polished moonstone set in a silver frame. It reveals hidden writing, invisible thresholds, and things that do not wish to be seen.
\item \textbf{Night-Bell:} A small bronze bell that rings only under moonlight. Its tone is soft—barely audible—but spirits hear it clearly.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Shadow-Cat:} A great cat made of darkness and stillness. It moves without sound. It can pass through a locked door if there is a gap—any gap—beneath it. It is most active under the full moon.
\item \textbf{Lunar Spirit:} A translucent figure of moonlight and mist. It can illuminate hidden thresholds, reveal creatures of shadow, and guide lost travelers. It cannot act—only point.
\item \textbf{The Moondrop Lens (Sentient):} A lens that speaks in whispers. It can reveal a hidden threshold (door, passage, weakness) once per night, but it will ask for a secret in return.
\end{itemize}

\subsection{Core Mechanic: The Moon's Phase}

The Moon Knight's power shifts with the moon's face. At the start of each session (or each in-game night), the GM determines the moon's phase:

\begin{itemize}
\item \textbf{New Moon:} +1 die to Stealth and Deception. The dark is kind.
\item \textbf{Crescent:} +1 die to Investigation and Insight. The light reveals.
\item \textbf{Half Moon:} +1 die to balance and resistance to being moved. The light and dark are equal.
\item \textbf{Gibbous:} +1 die to social rolls involving persuasion. The growing light lends confidence.
\item \textbf{Full Moon:} +2 dice to all physical actions. The silver light is power.
\end{itemize}

When the Moon Knight invokes a Rite under moonlight, they may reduce the Obligation cost by 1 (minimum 0). The moon witnesses.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Moonlit Mirror (4 XP)}

\textbf{Tags:} [TRUTH], [KNOWLEDGE], [REVELATION]

\textbf{Threshold:} A mirror or bowl of still water under moonlight.

Peer beyond surfaces. +1 die to detect illusions, veils, or hidden motives. Once per scene, ask one veiled question about the present; the GM answers in riddle or omen.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Crescent Whisper (5 XP)}

\textbf{Tags:} [COMMUNICATION], [BOND]

\textbf{Threshold:} A silvered thread knotted three times.

Deliver a secret thought to a single target; they hear it as if whispered beside them. Gain +1 die to aid their next action.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Triple Road (8 XP)}

\textbf{Tags:} [MOVEMENT], [GUIDANCE], [AREA]

\textbf{Threshold:} Three lit candles set at a crossroads.

Reveal a hidden path, door, or liminal way. Allies gain +2 dice to traverse or escape; enemies suffer -1 die to pursue.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Lunar Conjunction (7 XP)}

\textbf{Tags:} [SPIRIT], [KNOWLEDGE], [FATE]

\textbf{Threshold:} A ring of herbs burned beneath a dark moon.

Commune with veiled knowledge or subtle entities. Ask two questions; answers come in omen, dream-symbol, or paradox.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Moonbound Ascendant (14 XP)}

\textbf{Tags:} [TRANSFORM], [BOOST], [PROTECT]

\textbf{Threshold:} A cherished item surrendered under moonlight.

Enter a heightened lunar state. Choose two: +2 dice to Wits or Spirit, resist fear or beguilement, speak with unseen presences, or trace hidden connections with +2 Effect. Begin a 6-segment Moonbound clock.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Silver Memory (4 XP)}

\textbf{Requirement:} Lunara

You recall any event witnessed under moonlight with perfect clarity. Details normally lost to distraction or time remain sharp. This is a passive ability—and a burden.

\subsubsection{Quiet Step (6 XP)}

\textbf{Requirement:} Lunara

+1 die to stealth or subtle speech acts under moonlight. The silver light hides as much as it reveals.

\subsubsection{Seam-Walker (8 XP)}

\textbf{Requirement:} Lunara, Wits 3+

Once per scene, you may cross a minor threshold unseen—a door, a window, a gate, a bridge. Those watching must test Resolve (DV 4) to notice your passage. Mark 1 Fatigue after using this talent.

\subsection{Roleplaying a Moon Knight}

\begin{itemize}
\item You do not fight in the dark. You fight in the silver light. The moon is your witness, your ally, your judge.
\item You may be called to guide lost travelers, to hunt creatures that fear the light, or to protect those who walk the roads between.
\item You may struggle with the cycle. The moon wanes. Your power fades. The darkness is patient. You must learn to be patient too.
\item You may be asked to choose—between the light and the dark, between revelation and concealment, between the maiden, the mother, and the crone. The moon does not choose. It simply turns.
\end{itemize}

\subsection{Sample NPC: Watcher Sera of the Silver Path}

\textbf{Archetype:} Moon Knight

\textbf{Patron:} Lunara

\textbf{Codex:} A silver disc that glows under moonlight

\textbf{Thiasos:} A shadow-cat that moves without sound

\textbf{Personality:} Quiet, watchful, endlessly patient. She has walked the Ways Between for thirty years. She has never been lost. She has never been found.

\textbf{Conflict:} A child has wandered into the Green Host's territory on the night of the new moon. The moon is dark. The Knight's power is thin. She must choose: wait for the light, or walk into the dark alone.

\subsection{GM Guidance: Intrusions for Lunara}

\begin{itemize}
\item The moon is veiled by clouds. The Moon Knight's phase bonus is suspended until the clouds part—or until dawn.
\item A reflection in still water shows not the Moon Knight's face, but the face of someone they have lost. The water ripples. The face does not change.
\item The Moon Knight's silver disc grows cold. Something is blocking the moonlight—not clouds, but something else. Something hungry.
\item The Moon Knight dreams of walking under a moon that has three faces—maiden, mother, crone—all watching. When they wake, they cannot remember which face they saw last.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
