% −\textit{− coding: utf−8 −}−
% Chapter: Deck−Based Generators
% Comprehensive Resource Guide for Fate's Edge

\chapter{Deck−Based Generators}łabel{chap:deck_generators}

\section{Introduction to Deck Generators}łabel{sec:deckᵢntro}
∈dex{Deck Generators}

Deck generators in \textbf{Fate's Edge} transform random card draws into coherent narrative elements. Each deck has a distinct purpose and suit meanings so that randomness serves the story rather than derailing it. These generators provide structured inspiration for GMs while maintaining the game's narrative‑first philosophy.

\section{Standard Deck Structure}łabel{sec:deckₛtructure}
∈dex{Decks!Structure}

\textbf{Fate's Edge} uses several card‑based tools, each with specialised suit meanings:

\begin{description}
    ℹtem[\textbf{Travel Decks} (regional, 52‑card)] ∈dex{Decks!Travel}
        Used for journey content and location‑based adventures.
        \begin{itemize}
            ℹtem \textbf{Spade} = Place / Location
            ℹtem \textbf{Heart} = Actor / Faction
            ℹtem \textbf{Club} = Pressure / Complication
            ℹtem \textbf{Diamond} = Reward / Opportunity
        \end{itemize}
    ℹtem[\textbf{Deck of Consequences} (scene drama)] ∈dex{Decks!Consequences}
        Used for immediate complications and narrative twists during gameplay.
        \begin{itemize}
            ℹtem \textbf{Hearts} = Social / Emotional Fallout
            ℹtem \textbf{Spades} = Harm / Escalation
            ℹtem \textbf{Clubs} = Material Cost / Resource Drain
            ℹtem \textbf{Diamonds} = Magical / Spiritual Disturbance
        \end{itemize}
\end{description}

\begin{tcolorbox}[title=Important, colback=red!5!white, colframe=red!75!black]
Never mix suit meanings across decks. When rules reference \emph{Spade / Club / Diamond}, they mean the \textbf{Travel Deck}. When they say \emph{Hearts / Spades / Clubs / Diamonds}, they mean the \textbf{Deck of Consequences}.
\end{tcolorbox}

\section{Rank Severity and Timer Size}łabel{sec:rankₛeverity}
∈dex{Decks!Rank Severity}
∈dex{Timers!Size Determination}

Card rank sets the size / significance of the primary Timer :

\begin{itemize}
    ℹtem \textbf{2−−5 (Minor):}  4‑segment Timer ∈dex{Timers!4‑segment}
    ℹtem \textbf{6−−10 (Standard):} 6‑segment Timer ∈dex{Timers!6‑segment}
    ℹtem \textbf{J, Q, K (Major):} 8‑segment Timer ∈dex{Timers!8‑segment}
    ℹtem \textbf{Ace (Pivotal):} 10‑segment Timer ∈dex{Timers!10‑segment}
\end{itemize}

\noindent \textbf{Colour Influence:}
\begin{itemize}
    ℹtem \textbf{Black suits} (\textcolor{black}{$♠$}, \textcolor{black}{$♣$}): Travel hazards, tangible threats, fatigue. ∈dex{Decks!Black Suits}
    ℹtem \textbf{Red suits} (\textcolor{red}{$♥$}, \textcolor{red}{$♦$}): Social intrigue, reputational pressure, emotional complications. ∈dex{Decks!Red Suits}
\end{itemize}

\section{Draw Procedures}łabel{sec:drawₚrocedures}
∈dex{Decks!Draw Procedures}

\subsection{Quick Hook (2 cards)}łabel{subsec:quickₕook}
∈dex{Decks!Quick Hook}
Ideal for spontaneous scene generation or when players zag unexpectedly:
\begin{enumerate}
    ℹtem Draw one \textbf{Spade} (place) and one \textbf{Heart} (actor / faction).
    ℹtem Use the higher rank to set Timer size.
    ℹtem Combine elements into a simple, compelling scenario.
\end{enumerate}

\subsection{Full Seed (4 cards)}łabel{subsec:fullₛeed}
∈dex{Decks!Full Seed}
For full adventures or significant arcs:
\begin{enumerate}
    ℹtem Draw until one card of each suit appears:
        \begin{itemize}
            ℹtem \textbf{Spade} = Primary location ∈dex{Decks!Spade}
            ℹtem \textbf{Heart} = Main actor / faction ∈dex{Decks!Heart}
            ℹtem \textbf{Club} = Central complication ∈dex{Decks!Club}
            ℹtem \textbf{Diamond} = Key reward / opportunity ∈dex{Decks!Diamond}
        \end{itemize}
    ℹtem The highest rank sets the main Timer size.
    ℹtem If multiple face cards or Aces appear, create parallel Timers for secondary threats or opportunities.
\end{enumerate}

\subsection{Act Builder}łabel{subsec:act_builder}
∈dex{Decks!Act Builder}
Structure sessions or multi‑part adventures:
\begin{enumerate}
    ℹtem Draw three cards: setting (\textbf{Spade}), actor (\textbf{Heart}), complication (\textbf{Club}).
    ℹtem Treat \textbf{Diamond} cards drawn during play as foreshadowed opportunities or act payoffs.
    ℹtem Highest rank determines the session's primary challenge scope.
\end{enumerate}

\section{Using the Deck in Play}łabel{sec:deckᵢnₚlay}
∈dex{Decks!Usage}

\begin{enumerate}
    ℹtem Players roll; each die showing \textbf{1} generates \textbf{1 Story Beat (SB)}. ∈dex{Story Beats}
    ℹtem The GM chooses one method for that roll:
        \begin{enumerate}
            ℹtem \textbf{Direct Spend:} Translate SB into immediate consequences or timer ticks.
            ℹtem \textbf{Deck Draw:} Draw up to \textbf{min(SB, 3)} cards and synthesise a single twist guided by suit and highest rank.
        \end{enumerate}
    ℹtem Interpret the cards to create a coherent complication that advances the narrative.
\end{enumerate}

\section{Combo Rules}łabel{sec:comboᵣules}
∈dex{Decks!Combo Rules}
Special combinations add texture:

\begin{description}
    ℹtem[\textbf{Pair} (same rank)] Recurring motif with a twist. ∈dex{Decks!Pairs}
    ℹtem[\textbf{Run} (3+ sequential ranks)] Momentum −− reduce the main Timer by 1 segment. ∈dex{Decks!Runs}
    ℹtem[\textbf{Flush} (3+ same suit)] Strongly theme the act toward that suit's axis. ∈dex{Decks!Flushes}
    ℹtem[\textbf{Face + Ace}] Reveal a hidden patron or power behind the element. ∈dex{Decks!Face + Ace}
    ℹtem[\textbf{All one colour}] GM gains \textbf{+1 SB} to use in that scene. ∈dex{Decks!Colour Combos}
\end{description}

\section{Regional Generator Summary}łabel{sec:regionalₛummary}
∈dex{Decks!Regional Summary}

\begin{longtable}{p{3.3cm}p{5.3cm}p{5.0cm}}
∩tion{Regional Generator Summary}łabel{tab:regionalₛummary}\\
⊤rule
\textbf{Region} & \textbf{Theme} & \textbf{Special Mechanics} \\
∣rule
\endfirsthead
μlticolumn{3}{c}%
{{\bfseries \tablename\ θble{} −− continued}} \\
⊤rule
\textbf{Region} & \textbf{Theme} & \textbf{Special Mechanics} \\
∣rule
\endhead
⊥tomrule
\endfoot
Acasia & Broken Marches & Curse motifs; every Ace adds a lingering omen. \\
Aelaerem & Hearth & Hollow & Red‑thread motifs; Ace echoes quiet bells / watch‑geese. \\
Aeler & Crowns & Under‑Vaults & Stone / breath motifs; Ace: keys click, bells answer. \\
Aelinnel & Stone, Bough, Bright Things & Moonlight motifs; Ace adds a shortcut where none should be. \\
Black Banners & Condotta & Crowns & War & winter motifs; Ace: weapons remember, ice holds the dead. \\
Ecktoria & Marble & Fire & Imperial forms; Ace carves precedent in marble. \\
Kahfagia & Pilot's Mirror & Lantern‑law jurisdiction shifts; Ace redefines lanes. \\
Linn & Skerries & Storm‑Oaths & Sea omens; Ace: horns on wind, white horses on swell. \\
Mistlands & Bells, Salt, Breath & Breath / boundary motifs; Ace: bells answer across water. \\
Silkstrand & City of Bridges & Dyewater & Dye / bridge motifs; Ace adds a lingering omen. \\
Theona & Three Greens, No Ninth & "No Ninth" custom; Ace adds a telling omission. \\
Thepyrgos & City of a Thousand Stairs & Height / sound motifs; Ace echoes bells / wind / stair‑steps. \\
Ubral & Stone Between Spears & Upland motifs; Ace echoes horns / heather / stone. \\
Valewood & Empire Under Leaves & Empire echoes (J/Q/K add relic‑logic); Ace rearranges approach. \\
Vhasia & Fractured Sun & Broken‑sun motifs; Ace blots medal / scratches milestone. \\
Vilikari & Laurels & Longhouses & Two‑laws motifs; Ace shows wolf / eagle side‑by‑side. \\
Viterra & Last Kingdom & Legacy, parishes, and final‑stand themes. \\
Wilds & Roads, Ruins, Weather & Reskin palette for any biome. \\
Zakov & Salt & Serpent & Salt & serpent omens; Ace: tides remember, reefs shift, deep listens. \\
\end{longtable}

\section{NPC Generation Deck}łabel{sec:npc_generation}
∈dex{NPCs!Generation Deck}

Every NPC should feel like a person with desires, convictions, and contradictions. This deck lets you assemble a complete profile quickly by drawing one element from each category.

\subsection{Generation Categories}łabel{subsec:npc_categories}

\begin{description}
    ℹtem[\textbf{Ambition}] What they seek to achieve or obtain. ∈dex{NPCs!Ambition}
    ℹtem[\textbf{Belief}] The principle or philosophy guiding their worldview. ∈dex{NPCs!Belief}
    ℹtem[\textbf{Attitude}] How they present themselves and interact day‑to‑day. ∈dex{NPCs!Attitude}
    ℹtem[\textbf{Twist}] A contradiction or hidden facet that creates tension. ∈dex{NPCs!Twist}
\end{description}

\subsection{Using the NPC Generator}łabel{subsec:npcᵤsage}
Select or draw one from each column and consider the frictions between public ambition, private belief, surface attitude, and the twist.

\begin{longtable}{p{3.3cm}p{3.3cm}p{3.3cm}p{3.3cm}}
∩tion{NPC Generation Categories}łabel{tab:npc_categories}\\
⊤rule
\textbf{Ambition} & \textbf{Belief} & \textbf{Attitude} & \textbf{Twist} \\
∣rule
\endfirsthead
μlticolumn{4}{c}%
{{\bfseries \tablename\ θble{} −− continued}} \\
⊤rule
\textbf{Ambition} & \textbf{Belief} & \textbf{Attitude} & \textbf{Twist} \\
∣rule
\endhead
⊥tomrule
\endfoot
Power     & Might makes right       & Arrogant      & Secretly insecure \\
Wealth    & Ends justify means     & Charismatic   & Betraying their allies \\
Revenge   & Honour above all       & Cold          & Working for their enemy \\
Love      & Truth is sacred        & Friendly      & Hiding a dark past \\
Knowledge & Loyalty is paramount   & Paranoid      & Actually an impostor \\
Survival  & Family above all       & Cruel         & Deeply compassionate \\
Fame      & Justice must prevail  & Pious         & Corrupted by power \\
Freedom   & Fate can be changed    & Optimistic    & Hopelessly cynical \\
Protection& Tradition must be upheld & Pessimistic & Revolutionary at heart \\
Control   & Change is necessary    & Calculating   & Acts on impulse \\
Recognition& The system works      & Naive         & Cynical manipulator \\
\end{longtable}

\section{Practical Deck Usage Examples}łabel{sec:deckₑxamples}
∈dex{Decks!Examples}

\subsection{Example 1: Quick Scene Generation}łabel{subsec:example_quick}
The party detours through the Mistlands. The GM draws:
\begin{itemize}
    ℹtem \textbf{Spade (8)}: Ancient standing stones covered in moss.
    ℹtem \textbf{Heart (Queen)}: A territorial spirit guardian.
\end{itemize}
A 6‑segment Timer \emph{Spirit's Wrath} begins: the guardian demands tribute for safe passage.

\subsection{Example 2: Consequences During Play}łabel{subsec:example_consequences}
Alexander misses a stealth roll and generates \textbf{2 SB}. The GM draws:
\begin{itemize}
    ℹtem \textbf{Hearts (7)}: Social complication.
    ℹtem \textbf{Clubs (3)}: Resource cost.
\end{itemize}
Synthesis: \emph{A ceremonial urn shatters; cultists recognise your patron's mark. Future dealings will demand extra tribute and materials.}

\subsection{Example 3: NPC Creation}łabel{subsec:exampleₙpc}
Merchant in Valewood:
\begin{itemize}
    ℹtem \textbf{Ambition}: Wealth
    ℹtem \textbf{Belief}: Family above all
    ℹtem \textbf{Attitude}: Charismatic
    ℹtem \textbf{Twist}: Secretly compassionate
\end{itemize}
Result: \emph{A charming hard‑bargainer who supports a large family and quietly donates to orphanages −− even aiding struggling competitors.}

\section{GM Guidance for Deck Usage}łabel{sec:deck_guidance}
∈dex{Decks!GM Guidance}

\subsection{When to Use Which Deck}łabel{subsec:deckₛelection}
\begin{itemize}
    ℹtem \textbf{Travel Decks}: journey planning, location adventures, regional exploration.
    ℹtem \textbf{Deck of Consequences}: immediate twists during active scenes.
    ℹtem \textbf{NPC Generator}: fast creation with built‑in tension and hooks.
\end{itemize}

\subsection{Interpreting Card Draws}łabel{subsec:cardᵢnterpretation}
\begin{itemize}
    ℹtem Prioritise narrative coherence over literalism.
    ℹtem Use suits as inspiration, not constraints.
    ℹtem Combine cards into layered complications rather than parallel noise.
    ℹtem Remember: players can mitigate, pivot, or overcome deck outcomes.
\end{itemize}

\subsection{Balancing Randomness and Narrative}łabel{subsec:balancingᵣandomness}
\begin{itemize}
    ℹtem Draw when you want surprise or need a nudge.
    ℹtem Ignore or modify draws that don"t serve the current story.
    ℹtem Treat combinations as creative prompts, not mandates.
    ℹtem The goal is to \emph{enhance} the narrative, not derail it.
\end{itemize}

% End of chapter