% -*- coding: utf-8 -*-
% Chapter: Travel Reference
% Comprehensive Resource Guide for Fate's Edge

\chapter{Travel Reference}\label{chap:travel_reference}

\section{Introduction to Travel}\label{sec:travel_intro}
\index{Travel}\index{Decks!Travel}\index{Deck Generators}

Travel in \textbf{Fate's Edge} is not just movement between locations—it is an opportunity for adventure, discovery, and complication. The travel system uses deck‑based generators to create meaningful journeys that advance the story while presenting challenges and opportunities.\index{Journeys}\index{Exploration}

\section{Core Travel Procedure}\label{sec:core_travel}
\index{Travel!Core procedure}\index{Procedures!Travel}

For each leg of a journey, follow this structured approach.\index{Legs of travel}

\subsection{Card Draw Sequence}\label{subsec:card_draw}
\index{Travel!Card draw}\index{Decks!Travel!Draw sequence}

Draw 3–4 cards using the appropriate regional decks:

\begin{description}
    \item[\textbf{Spade} from the destination deck] Sets the scene (place and environment).\index{Travel!Spade}\index{Suits!Spade}\index{Places and terrain}
    \item[\textbf{Heart} from the destination deck] Introduces the local actor or faction.\index{Travel!Heart}\index{Suits!Heart}\index{Factions}
    \item[\textbf{Club} from the Wilds (general hazards) or destination (if strongly policed)] Brings pressure and complications.\index{Travel!Club}\index{Suits!Club}\index{Hazards}
    \item[\textbf{Diamond} from the authority that controls the route] Represents papers, escorts, rights, or exceptions needed.\index{Travel!Diamond}\index{Suits!Diamond}\index{Permits and papers}
\end{description}

\subsection{Travel Clock Setup}\label{subsec:travel_clock}
\index{Travel!Clocks}\index{Clocks!Travel}

Set a travel clock based on the highest card rank drawn:

\begin{longtable}{c c p{8cm}}
\caption{Travel Clock Sizes}\label{tab:travel_clock_sizes}\\
\toprule
\textbf{Rank} & \textbf{Clock Size} & \textbf{Typical Journey} \\
\midrule
2–5   & 4 segments & Short, straightforward journey. \\
6–10  & 6 segments & Standard journey. \\
J, Q, K & 8 segments & Extended or complex journey. \\
Ace   & 10 segments & Epic or highly dangerous journey. \\
\bottomrule
\end{longtable}

\subsection{Resolution and Consequences}\label{subsec:travel_resolution}
\index{Travel!Consequences}\index{Consequences!Travel}

\begin{itemize}
    \item On success: advance to the next leg of the journey.\index{Success!Travel}
    \item On failure: mark delay, debt, or diversion and resolve consequences in the fiction.\index{Failure!Travel}
    \item Each segment filled typically represents significant progress or a meaningful encounter.\index{Encounters!Travel}
\end{itemize}

\section{Travel Modes and Special Rules}\label{sec:travel_modes}
\index{Travel!Mode rules}\index{Modes of travel}

\subsection{Sea Travel}\label{subsec:sea_travel}
\index{Travel!Sea legs}\index{Sea travel}\index{Maritime!Travel}

Sea journeys (Amaranthine / Dolmis / Aberderrin routes) follow special considerations:\index{Amaranthine}\index{Dolmis}\index{Aberderrin}

\begin{itemize}
    \item If Theona or Valewood 9s appear in the seed, add an omission or taboo to the leg.\index{Travel!Taboos}\index{Theona}\index{Valewood}
    \item Navigational challenges use \emph{Sailing} or \emph{Survival} skills with modified DVs.\index{Skills!Survival}\index{Skills!Sailing}\index{Difficulty Value (DV)}
    \item Weather and sea conditions act as persistent complications.\index{Hazards!Weather}\index{Sea states}
\end{itemize}

\subsection{Underground Passes (Aeler)}\label{subsec:underways}
\index{Travel!Underways}\index{Aeler!Underways}\index{Underground routes}

Aeler’s underground routes have unique mechanics:\index{Aeler}

\begin{itemize}
    \item Any Ace can convert a surface route to an under‑route.\index{Travel!Aces}\index{Aces}
    \item Diamond cards from Aeler always codify outcomes—no roll required—for sealed doors, breath time, and escorted segments.\index{Aeler!Diamonds}\index{Permits and papers!Aeler}
    \item Underground travel uses different skill sets (\emph{Engineering}, \emph{Geology}, \emph{Survival}).\index{Skills!Engineering}\index{Skills!Geology}\index{Skills!Survival}
\end{itemize}

\subsection{River Travel}\label{subsec:river_travel}
\index{Travel!Rivers}\index{River travel}

River journeys present distinct challenges:

\begin{itemize}
    \item \textbf{Ecktoria / Viterra rivers:} bridges, booms, and legal checkpoints.\index{Ecktoria}\index{Viterra}\index{Checkpoints}
    \item \textbf{Mistlands / Linn waters:} reed‑mazes and bell‑line navigation systems.\index{Mistlands}\index{Linn}\index{Bell-lines}
    \item Current strength and river conditions affect travel pace and difficulty.\index{Currents}\index{Difficulty Value (DV)!River}
\end{itemize}

\subsection{Frontier and Border Crossings}\label{subsec:frontier_travel}
\index{Travel!Frontier blends}\index{Borders}\index{Frontiers}

When traveling between regions with different laws or authorities:

\begin{itemize}
    \item Draw two Diamond cards (one from each controlling authority).\index{Permits and papers!Dual authority}
    \item Choose which authority’s laws you will be judged by at the journey’s end.\index{Jurisdiction}
    \item Border crossings typically add +2 segments to the travel clock.\index{Travel!Law boundaries}\index{Clocks!Travel!Modifiers}
\end{itemize}

\section{Major Regional Routes}\label{sec:regional_routes}
\index{Travel!Regional routes}\index{Routes!Major}

\subsection{Amaranthine Coastway}\label{subsec:amaranthine_coastway}
\index{Travel!Amaranthine Coastway}\index{Amaranthine}

\textbf{Route:} Kahfagia $\rightarrow$ Ecktoria $\rightarrow$ Acasia $\rightarrow$ Marcott (Vhasia) $\rightarrow$ Fairport (Viterra).\index{Kahfagia}\index{Acasia}\index{Vhasia!Marcott}\index{Viterra!Fairport}

\begin{itemize}
    \item Draw Spade+Heart from the port you are entering.\index{Ports}
    \item Draw Club from the Wilds deck (gales, quarantine, harbor booms).\index{Wilds}\index{Hazards!Gales}\index{Quarantine}
    \item Draw Diamond from the port’s deck (permits, berths), or Kahfagia if convoy / lantern law applies.\index{Lantern law}\index{Permits and papers!Ports}
    \item \textbf{Note:} Silkstrand is Acasia’s sole metropolis and a major trading hub.\index{Acasia!Silkstrand}\index{Trade hubs}
\end{itemize}

\subsection{Astroegro Straits}\label{subsec:astroegro_straits}
\index{Travel!Astroegro Straits}\index{Thepyrgos!Straits}\index{Straits}

Thepyrgos controls this crucial maritime hinge.\index{Thepyrgos}

\begin{itemize}
    \item Draw Spade+Heart and Club from Thepyrgos deck (stairs, boom, synod).\index{Boom chains}\index{Synod}
    \item Draw Diamond from Thepyrgos (harbor pass, crane allotment).\index{Permits and papers!Harbor}
    \item Switch to Dolmis / Amaranthine modules once the boom is lowered.\index{Dolmis}\index{Amaranthine}
\end{itemize}

\subsection{Dolmis Sea Circuits}\label{subsec:dolmis_circuits}
\index{Travel!Dolmis circuits}\index{Sea circuits}

\textbf{Route:} Fairport (Viterra) $\rightarrow$ Theona (Three Greens) $\rightarrow$ Ubral fjords $\rightarrow$ Aelinnel west shore.\index{Theona}\index{Ubral}\index{Aelinnel}

\begin{itemize}
    \item Spade+Heart from destination deck.\index{Suits!Spade}\index{Suits!Heart}
    \item Club from Linn or Wilds decks.\index{Linn}\index{Wilds}
    \item Diamond from Theona when island custom matters.\index{Theona!Dolmis}\index{Island customs}
    \item Island‑hopping presents unique navigation challenges.\index{Navigation!Islands}
\end{itemize}

\subsection{Aelerian Underground Passes}\label{subsec:aelerian_passes}
\index{Travel!Aelerian passes}\index{Aeler!Passes}

\textbf{Route:} Vhasia / Viterra / Ubral south slopes $\rightarrow$ Aeler gates $\rightarrow$ Mistlands.\index{Vhasia}\index{Viterra}\index{Ubral}\index{Aeler!Gates}\index{Mistlands}

\begin{itemize}
    \item Spade+Heart and Club from Aeler deck.\index{Aeler}
    \item Diamond from Aeler (Underway Pass, Key‑Writ, Breath Time).\index{Aeler!Gates}\index{Permits and papers!Aeler}
    \item Underground routes bypass surface dangers but present their own risks.\index{Underground routes!Risks}
\end{itemize}

\subsection{Shadow Corridors (Valewood)}\label{subsec:shadow_corridors}
\index{Travel!Shadow corridors}\index{Valewood!Shadow corridors}

Thin Shore route along Valewood’s east coast toward Zakov.\index{Thin Shore}\index{Zakov}

\begin{itemize}
    \item Spade from Valewood deck.\index{Valewood}
    \item Heart from Mistlands or Valewood deck.\index{Mistlands}
    \item Club from Mistlands deck (wraiths, bell‑line failures).\index{Wraiths}\index{Bell-lines}
    \item Diamond from either Mistlands (Ward‑salt, Lantern Writ) or Valewood (Way‑cord, Truce‑bough).\index{Ward-salt}\index{Lantern writ}\index{Way-cord}\index{Truce-bough}
\end{itemize}

\begin{tcolorbox}[title=Rule of 9s, colback=red!5!white, colframe=red!75!black, sharp corners]
    Any \textbf{9} in the seed adds an omission or taboo: a missing step, an unsaid name, an unseen guest.\index{Travel!Rule of 9s}
\end{tcolorbox}

\subsection{Green Gate Tidal Routes (Aelinnel)}\label{subsec:green_gate}
\index{Travel!Green Gate}\index{Aelinnel!Green Gate}\index{Tides}

Aelinnel’s tide‑dependent passages:

\begin{itemize}
    \item Spade+Club from Aelinnel deck (tides, geasa).\index{Geasa}
    \item Diamond from Aelinnel (Tide‑path Key, Host‑rights).\index{Host-rights}\index{Permits and papers!Aelinnel}
    \item Timing is critical—miss the tide and wait for the next cycle.\index{Timing!Tides}
\end{itemize}

\subsection{Major River Systems}\label{subsec:river_systems}
\index{Travel!River roads}\index{Rivers!Major systems}

\begin{description}
    \item[\textbf{Belworth River}] Boundary between Vhasia and Viterra.\index{Belworth River}\index{Vhasia}\index{Viterra}
        \begin{itemize}
            \item Twin ports at the mouth (Marcott and Fairport).\index{Vhasia!Marcott}\index{Viterra!Fairport}
            \item Tarlington Falls requires lift / portage to plateau towns.\index{Tarlington Falls}
            \item Use Viterra for east‑bank audits, Vhasia for west‑bank politics.\index{Audits}\index{Politics!Vhasia}
        \end{itemize}
    \item[\textbf{Yloka River}] Northern raid routes into Aberderrin and Dolmis.\index{Yloka}\index{Aberderrin}\index{Dolmis}
        \begin{itemize}
            \item Overlay Linn Clubs for hazards.\index{Linn}\index{Hazards!Linn}
            \item Draw destination for Spade+Heart.\index{Suits!Spade}\index{Suits!Heart}
        \end{itemize}
    \item[\textbf{Acasian Rivers}] Hazardous approaches to Silkstrand.\index{Acasia!Rivers}\index{Acasia!Silkstrand}
        \begin{itemize}
            \item Use Acasia Clubs (feuds / levies / curse).\index{Feuds}\index{Levies}\index{Curses}
            \item Diamonds represent condotta permits, bridge rights.\index{Condotta}\index{Bridge rights}
        \end{itemize}
\end{description}

\subsection{Steppe Frontier Routes}\label{subsec:steppe_frontiers}
\index{Travel!Steppe frontiers}\index{Ykrul}\index{Vilikari}

Ykrul $\leftrightarrow$ Vilikari $\leftrightarrow$ Ecktoria / Acasia borders:\index{Ecktoria}\index{Acasia}

\begin{itemize}
    \item Spade from the road (Wolf Road or Foedus Stone).\index{Wolf Road}\index{Foedus Stone}
    \item Heart from the counterpart region (envoys, comitatus).\index{Envoys}\index{Comitatus}
    \item Club from Wilds or frontier decks (rasputitsa, hostage protocol).\index{Rasputitsa}\index{Hostage protocol}
    \item Diamond from either Ykrul (paiza, remounts) or Vilikari (foedus seal).\index{Paiza}\index{Remounts}\index{Foedus}
\end{itemize}

\section{Strategic Gateways and Control Points}\label{sec:gateways}
\index{Travel!Gateways}\index{Control points}

\begin{longtable}{p{3cm}p{2.8cm}p{3.2cm}p{3.8cm}}
\caption{Major Travel Gateways and Control Points}\label{tab:gateways}\\
\toprule
\textbf{Gateway} & \textbf{Controlling Deck} & \textbf{Diamond Source} & \textbf{Typical Requirements} \\
\midrule
\endfirsthead
\multicolumn{4}{c}%
{{\bfseries \tablename\ \thetable{} – continued}} \\
\toprule
\textbf{Gateway} & \textbf{Controlling Deck} & \textbf{Diamond Source} & \textbf{Typical Requirements} \\
\midrule
\endhead
\bottomrule
\endfoot
Kassamira (Kahfagia) & Kahfagia & Kahfagia & Convoy letter; lantern‑law warrant; pilotage token \\
Ecktoria Capital Port & Ecktoria & Ecktoria & Berth priority; coin‑house credit; procession permit \\
Silkstrand & Acasia & Acasia & Exchange floor access; condotta permit; watergate priority \\
Marcott & Vhasia & Vhasia & Letters patent; safe‑conduct; fairground precedence \\
Fairport & Viterra & Viterra & Customs seal; ferry priority; Dawn escort \\
Tarlington Falls & Viterra & Viterra & Lift priority; bridge/portage pass; County ruling \\
Thepyrgos Boom & Thepyrgos & Thepyrgos & Harbor pass; crane allotment; Synod indulgence \\
High‑Mist Pass & Aeler & Aeler & Underway pass; Key‑Writ; breath‑time allotment \\
Payden’s Port & Mistlands & Mistlands & Lantern writ; ward‑salt; wraith indemnity \\
Thin Shore & Valewood & Valewood / Mistlands & Way‑cord; dew‑mirror; lantern writ \\
Green Gate & Aelinnel & Aelinnel & Tide‑path key; host‑right bough; salt license \\
Ubral Passes & Ubral & Aeler / Ubral & Vurim pass‑ring; hill‑fort shelter writ; guide’s braid \\
Thing‑holm & Linn & Linn & Thing ruling; harbor‑green; escort writ \\
Foedus Stone & Vilikari & Vilikari / Ykrul & Foedus renewal; remount allotment; hostage treaty \\
Zakov Roadstead & Theona / Linn & Theona / Linn & Raid‑truce ribbon; whale‑road escort; market immunity \\
\end{longtable}

\section{Special Travel Rules and Regional Features}\label{sec:special_rules}
\index{Travel!Special rules}\index{Regional features}

\subsection{Theona and Valewood Taboos}\label{subsec:theona_taboos}
\index{Travel!Taboos}\index{Theona!Taboos}\index{Valewood!Taboos}

Whenever a 9 appears in a travel seed involving Theona or Valewood:
\begin{itemize}
    \item Add an omission: missing step, unsaid name, unseen guest.\index{Omissions}
    \item If the 9 is a Diamond, you may break the taboo once—but someone will come to collect.\index{Suits!Diamond}\index{Debts}
    \item These taboos create ongoing narrative hooks and complications.\index{Hooks}\index{Complications!Taboos}
\end{itemize}

\subsection{Aeler Ace Route Manipulation}\label{subsec:aeler_aces}
\index{Travel!Aces}\index{Aeler!Aces}

Aeler’s underground network allows special route manipulation:
\begin{itemize}
    \item An Ace can rewrite routes beneath the mountains.\index{Aces!Route rewrite}
    \item You may swap your Spade to Aeler (keeping other seed cards) and continue under‑vault procedures.\index{Suits!Spade}\index{Aeler!Underways}
    \item This represents discovering or activating hidden passages.\index{Hidden passages}
\end{itemize}

\subsection{Echoing Omens and Motifs}\label{subsec:echoing_omens}
\index{Travel!Echoing omens}\index{Omens}\index{Motifs}

Any Ace drawn during travel adds a recurring motif:
\begin{itemize}
    \item Examples: ash on the wind, a bell heard underground, a landmark that reappears.\index{Ash omens}\index{Bells}\index{Landmarks}
    \item These motifs can be referenced later for narrative advantage or complication.\index{Foreshadowing}
    \item Makes the world feel interconnected and memorable.\index{Worldbuilding}
\end{itemize}

\section{Worked Travel Examples}\label{sec:worked_examples}
\index{Travel!Worked examples}\index{Examples!Travel}

\subsection{West–to–East Coastal Journey}\label{subsec:coastal_journey}
\index{Travel!Coastal haul}\index{Kassamira}\index{Fairport}

\textbf{Leg 1: Kassamira $\rightarrow$ Ecktoria}
\begin{itemize}
    \item Seed: Ecktoria Spade (Aqueduct arcades), Ecktoria Heart (Coin‑house factor), Wilds Club (gale), Kahfagia Diamond (convoy letter).\index{Aqueducts}\index{Coin-houses}\index{Hazards!Gales}\index{Convoys}
    \item Clock: 6 segments.\index{Clocks!6-segment}
    \item \textit{Narrative:} Cross under coin‑house eyes with convoy priority; bad weather strains schedules.
\end{itemize}

\textbf{Leg 2: Ecktoria $\rightarrow$ Silkstrand}\index{Silkstrand!Approach}
\begin{itemize}
    \item Seed: Acasia Spade (Three‑Queens Bridge), Acasia Heart (Dyers’ Guildmistress), Acasia Club (loom strike), Acasia Diamond (Exchange pass).\index{Three-Queens Bridge}\index{Guilds!Dyers}\index{Strikes}\index{Exchange passes}
    \item Clock: 6–8 segments depending on unrest severity.\index{Clocks!6-segment}\index{Clocks!8-segment}
\end{itemize}

\textbf{Leg 3: Silkstrand $\rightarrow$ Marcott}\index{Marcott!Approach}
\begin{itemize}
    \item Seed: Vhasia Spade (Pont‑du‑Tithe), Vhasia Heart (Parlement clerk), Vhasia Club (coin rumor), Vhasia Diamond (letters patent).\index{Pont-du-Tithe}\index{Parlement}\index{Rumors}\index{Letters patent}
\end{itemize}

\textbf{Leg 4: Marcott $\rightarrow$ Fairport}\index{Fairport!Approach}
\begin{itemize}
    \item Seed: Viterra Spade (Fairport tideworks), Viterra Heart (shipwright), Linn Club (boom lifts), Viterra Diamond (customs seal).\index{Tideworks}\index{Shipwrights}\index{Boom lifts}\index{Customs seal}
    \item Clock: 6 segments.\index{Clocks!6-segment}
\end{itemize}

\subsection{Mountain Pass Expedition}\label{subsec:mountain_pass}
\index{Travel!Mountain passes}\index{Expeditions}

\textbf{Leg 1: Silkstrand $\rightarrow$ Aeler Gate}\index{Aeler!Gate approach}
\begin{itemize}
    \item Seed: Aeler Spade (Avalanche gallery), Aeler Heart (Geometer), Aeler Club (Engineer requisition), Aeler Diamond (Underway Pass).\index{Avalanche galleries}\index{Geometers}\index{Engineers}\index{Underway Pass}
    \item Underground route bypasses surface dangers.
\end{itemize}

\textbf{Leg 2: Aeler Gate $\rightarrow$ Mistlands}\index{Mistlands!Approach}
\begin{itemize}
    \item Seed: Mistlands Spade (Bell‑Line levee), Mistlands Heart (Bell‑warden), Mistlands Club (wraith crossing), Mistlands Diamond (Ward‑salt).\index{Bell-wardens}\index{Levees}\index{Wraiths}\index{Ward-salt}
    \item Mistlands present unique navigation challenges.\index{Navigation!Mistlands}
\end{itemize}

\subsection{Shadow Route Travel}\label{subsec:shadow_route}
\index{Travel!Shadow routes}

\textbf{Leg 1: Payden’s Port $\rightarrow$ Thin Shore}\index{Thin Shore!Approach}\index{Payden's Port}
\begin{itemize}
    \item Seed: Valewood Spade (Green lane), Mistlands Heart (Protectorate clerk), Mistlands Club (bell failure), Mistlands Diamond (Lantern writ).\index{Green lanes}\index{Protectorate}\index{Bell-lines}\index{Lantern writ}
    \item Rule of 9s applies for additional complications.\index{Travel!Rule of 9s}
\end{itemize}

\textbf{Leg 2: Thin Shore Transit}\index{Thin Shore!Transit}
\begin{itemize}
    \item Seed: Valewood Spade (Unfound stile), Valewood Heart (Path‑warden), Valewood Club (Sweet wind), Valewood Diamond (Way‑cord).\index{Unfound stiles}\index{Path-wardens}\index{Sweet winds}\index{Way-cord}
    \item Shadow corridors require special navigation skills.\index{Navigation!Shadow corridors}
\end{itemize}

\subsection{Frontier Diplomacy Journey}\label{subsec:frontier_diplomacy}
\index{Travel!Frontier mediation}\index{Diplomacy!Frontier}

\textbf{Leg: Foedus Stone Parley}\index{Foedus Stone!Parley}
\begin{itemize}
    \item Seed: Vilikari Spade (Foedus Stone), Ykrul Heart (Noyan envoy), Wilds Club (rasputitsa), Vilikari Diamond (Foedus Renewal).\index{Vilikari}\index{Ykrul}\index{Noyan}\index{Foedus Renewal}
    \item Additional draw: Ykrul Diamond (Paiza) for border crossing rights.\index{Paiza}\index{Permits and papers!Ykrul}
    \item Choose which authority’s laws govern your passage.\index{Jurisdiction}
\end{itemize}

\section{Travel Pace and Consequence Guidelines}\label{sec:pace_consequences}
\index{Travel!Pace}\index{Pacing}

\subsection{Standard Travel Times}\label{subsec:travel_times}
\index{Travel!Time scales}

\begin{longtable}{p{4cm}p{10cm}}
\caption{Standard Travel Times}\label{tab:travel_times}\\
\toprule
\textbf{Terrain Type} & \textbf{Clock Size and Modifiers} \\
\midrule
Road / River leg & 6‑segment clock standard; +2 segments if crossing law boundaries. \\
Sea strait / Mountain pass & 8‑segment clock; –2 segments with priority documents. \\
Shadow corridor & 6‑segment clock; Ace adds free omen for future use. \\
Urban travel & 4‑segment clock; complications are more social than environmental. \\
\bottomrule
\end{longtable}

\subsection{Failure Consequences}\label{subsec:failure_consequences}
\index{Travel!Failure consequences}\index{Consequences!Failure}

Standard failure outcomes include:
\begin{description}
    \item[\textbf{Delay}] Lose time; enemies advance their plans.\index{Delay}
    \item[\textbf{Diversion}] Forced alternate route; draw new Club card for additional complications.\index{Diversion}\index{Suits!Club}
    \item[\textbf{Debt}] Promise, fine, or favour owed to the controlling authority.\index{Debt}
    \item[\textbf{Discovery}] Unexpected find that creates new opportunities or dangers.\index{Discovery}
\end{description}

\subsection{Success Benefits}\label{subsec:success_benefits}
\index{Travel!Success benefits}\index{Success!Travel}

Successful travel can provide:
\begin{itemize}
    \item Information about upcoming regions or threats.\index{Information!Regional}
    \item New contacts or faction relationships.\index{Factions!Contacts}
    \item Resource acquisition or trade opportunities.\index{Trade}\index{Resources}
    \item Narrative advantages for future scenes.\index{Positioning!Narrative}
\end{itemize}

\clearpage
\section{Locations Lore and Generators}
\input{generators/acasia.tex}
\clearpage
\input{generators/aelaerem.tex}
\clearpage
\input{generators/aeler.tex}
\clearpage
\input{generators/aelinnel.tex}
\clearpage
\input{generators/black_banner.tex}
\clearpage
\input{generators/dungeons.tex}
\clearpage
\input{generators/ecktoria.tex}
\clearpage
\input{generators/linns.tex}
\clearpage
\input{generators/midh-akhaz.tex}
\clearpage
\input{generators/mistlands.tex}
\clearpage
\input{generators/silkstrand.tex}
\clearpage
\input{generators/theona.tex}
\clearpage
\input{generators/thepyrgos.tex}
\clearpage
\input{generators/ubral.tex}
\clearpage
\input{generators/valewood.tex}
\clearpage
\input{generators/vhasia.tex}
\clearpage
\input{generators/vilikari.tex}
\clearpage
\input{generators/viterra.tex}
\clearpage
\input{generators/ways_between.tex}
\clearpage
\input{generators/wilderness.tex}
\clearpage
\input{generators/ykrul.tex}
\clearpage
\input{generators/zakov.tex}
\clearpage

\subsection{Politics of the Mortal Realms}\label{subsec:politics_mortal_realms}
\index{Politics}

Power in Fate's Edge is transactional, territorial, and personal. Every Patron, every city‑state, every clan has teeth. Characters do not merely travel through the world—they collide with the systems that run it.

\subsubsection{Patronage and Power Dynamics}\index{Patrons!Politics}

Patrons grant Rites and miracles, but demand service, obedience to theme, or the slow gravity of Corruption and Obligation. Runekeepers, Cantors, and Invokers navigate the politics of the divine as surely as nobles do on mortal thrones.

Rivalries are explicit:
\begin{itemize}
    \item Malachai vs. Oath of Flame \& Light: cursed blessing against purifying vow.
    \item Aveh vs. Varnek Karn: storm‑born freedom against cosmic accounting.
    \item Inquisitor Prime vs. Outsider Patrons: zealotry against the arcane.
\end{itemize}
Allying with one Patron often earns the enmity of another. A boon is never free; the world notices who you serve.

\subsubsection{Regional Governance and Conflict}\index{Regions!Politics}

Each realm runs by its own rules. To take power, keep power, or even survive, characters must learn how that region governs.

\begin{description}
    \item[Vhasia:] fractured monarchy, duelling claimants, and the Chevauchée as political warfare.
    \item[Vilikari:] Utaran law and steppe clan law—a case can be won in one court and lost in the other.
    \item[Aeler Holds:] guilds, lineages, and territorial rights; justice measured in stone, steel, and precedent.
    \item[Merchant Republics:] Kahfagia, Silkstrand, and Thepyrgos ruled by coin‑houses and guild votes.
    \item[Ykrul, Linn, Ubral:] clan justice, blood‑price, guest‑right, and the Thing or the Moot.
    \item[Theocracy:] Ecktoria’s Everflame and the Valewood Courts where scripture is law.
\end{description}
Politics means charters, hostage treaties, levies, blood feuds, Synod judgments, and the quiet power of written law.

\subsubsection{Institutional Power}\index{Institutions}

Thrones are not the only engines of authority.
\begin{itemize}
    \item Guilds control crafts, trade, and representation.
    \item Mercenary companies make or break wars (Black Banners, Condotta).
    \item Legal systems vary wildly: Utaran Code, Aeler Measure, Hedge‑Law, Synods.
    \item Magical colleges and Runekeeper orders police the arcane.
\end{itemize}
A crown can be ignored; a guild charter cannot.

\subsubsection{Resource Control and Economic Leverage}\index{Resources}

Salt, iron, grain, stone, glass, dye, and wine decide who rises and who starves. Control of bridges, ferries, straits, and toll rights is as valuable as any army. A dragon hoards gold. A duke hoards trade routes.

\subsubsection{Social Currency and Reputation}\index{Reputation}

Names have weight. Honour, geasa, rescue debts, hostage treaties, and guest‑right shape diplomacy as much as swords. Fame attracts allies—and rivals. A single broken oath can change a kingdom’s fate.

\paragraph{In Summary.}
Characters do not merely slay monsters. They enter the machinery of power: Patrons, guilds, courts, clans, and crowns. Every choice matters. Every consequence echoes.

% End of chapter
