% -*- coding: utf-8 -*-
% Chapter: Resource Management
% Comprehensive Resource Guide for Fate's Edge

\chapter{Resource Management}\label{chap:resource_management}
\index{Resources}\index{Resource Management}

\section{Introduction}\label{sec:resource_intro}
\index{Resources!Overview}

\indexterm{Resource Management} in \textbf{Fate's Edge} focuses on tracking the tangible and intangible assets that enable character actions while creating narrative tension. Rather than micromanaging every item, the system uses abstract clocks and conditions that trigger when dramatically appropriate. This keeps the focus on story consequences rather than bookkeeping.\index{Clocks}\index{Conditions}

\section{Supply Clock}\label{sec:supply_clock}
\index{Supply}\index{Resources!Supply Clock}

The \indexterm{Supply Clock} is a shared condition for the entire party, representing food, water, ammunition, and basic gear. It tracks the group’s overall readiness for extended endeavours.\index{Party resources}

\subsection{Supply States}\label{subsec:supply_states}
\index{Supply!States}\index{Clocks!Supply}

\begin{longtable}{p{3cm} p{10cm}}
\caption{Supply Clock States}\label{tab:supply_states}\\
\toprule
\textbf{State} & \textbf{Effect} \\
\midrule
\endfirsthead
\multicolumn{2}{c}%
{{\bfseries \tablename\ \thetable{} – continued}} \\
\toprule
\textbf{State} & \textbf{Effect} \\
\midrule
\endhead
\bottomrule
\endfoot
Full Supply (0 segments) & The party is well‑equipped and prepared. No penalties. \\
Low Supply (2 segments) & Minor narrative complications: bland food, damaged arrows, thinning waterskins, worn gear. \\
Dangerously Low (3 segments) & Each character gains \textbf{Fatigue 1} due to exhaustion and deprivation. \\
Out of Supply (4 segments) & Severe penalties apply; characters face starvation, dehydration, and failing gear with significant mechanical consequences. \\
\end{longtable}

\subsection{Filling the Supply Clock}\label{subsec:filling_supply}
\index{Supply!Filling}\index{Story Beats}

The Supply Clock advances under the following circumstances:
\begin{itemize}
    \item Harsh travel conditions or lost pack animals (GM fiat).\index{GM fiat}
    \item GM spends \textbf{2+ Story Beats (SB)} to represent resource depletion.\index{Story Beats!Spends}
    \item The party chooses to travel light for speed or stealth advantages.\index{Travel!Travel light}
    \item Failed foraging or resource‑gathering attempts.\index{Survival!Foraging}
\end{itemize}

\subsection{Emptying the Supply Clock}\label{subsec:emptying_supply}
\index{Supply!Emptying}\index{Downtime}

The Supply Clock can be reduced through:
\begin{itemize}
    \item Reaching civilisation or safe haven resets to \textbf{Full Supply}.\index{Supply!Civilisation}
    \item Successful foraging / hunting: group \emph{Survival} check (DV 2) clears 1 segment.\index{Survival!Foraging}\index{Difficulty Value}
    \item Downtime spent in relative safety removes 1 segment.\index{Downtime!Supply}
    \item Purchasing or trading for supplies in settlements.\index{Supply!Purchasing}
\end{itemize}

\section{Fatigue}\label{sec:fatigue}
\index{Fatigue}\index{Conditions!Fatigue}

\indexterm{Fatigue} represents physical exhaustion, hunger, emotional strain, and spiritual depletion. It accumulates through extended effort, deprivation, or magical backlash.\index{Backlash}

\subsection{Clearing Fatigue}\label{subsec:clearing_fatigue}
\index{Fatigue!Clearing}\index{Rest}

\begin{itemize}
    \item A night’s rest with adequate Supply removes 1 level of Fatigue.
    \item You cannot clear Fatigue if the party’s Supply clock is \emph{Dangerously Low} or \emph{Out of Supply}.\index{Supply}
    \item Magical healing or special abilities may provide additional Fatigue recovery.\index{Magic!Healing}
    \item Extended downtime (3+ days) in safety clears all Fatigue.\index{Downtime!Fatigue recovery}
\end{itemize}

\section{Gear Condition}\label{sec:gear_condition}
\index{Gear}\index{Resources!Gear condition}

Gear in \textbf{Fate's Edge} does not have hit points or detailed durability tracking. Instead, equipment suffers consequences only when drama demands it.\index{Narrative-first}

\subsection{Compromised Items}\label{subsec:compromised_gear}
\index{Gear!Compromised}\index{Story Beats}

\begin{itemize}
    \item Compromised status is introduced via SB spends or narrative consequence.\index{Story Beats!Complications}
    \item A \textbf{Compromised} item gives \(-1\) die on relevant rolls until repaired.\index{Penalties}
    \item Multiple Compromised items affecting the \emph{same} action do not stack penalties.
\end{itemize}

\subsection{Breaking Point}\label{subsec:gear_breaking}
\index{Gear!Breaking}

If a Compromised item suffers another significant setback, it breaks entirely and becomes unusable.

\subsection{Repair Options}\label{subsec:gear_repair}
\index{Gear!Repair}\index{Craft}

\begin{description}
    \item[\indexterm{Field Repair}] Temporary fix; requires \emph{Craft} or \emph{Survival} check (DV 2) to remove the penalty for one scene.\index{Survival}\index{Difficulty Value}
    \item[\indexterm{Proper Repair}] Permanent restoration; requires proper tools, materials, and significant downtime.\index{Downtime}
    \item[\indexterm{Magical Repair}] Certain spells or rituals can instantly restore gear, but may carry Obligation or Backlash costs.\index{Obligation}\index{Backlash}
\end{description}

\subsection{Equipment Enchantments Quick Reference}\label{subsec:enchantments}
\index{Enchantments}

\begin{longtable}{p{4cm} p{9cm}}
\caption{Equipment Enchantment Guidelines}\label{tab:enchantments}\\
\toprule
\textbf{Type} & \textbf{Cost} \\
\midrule
Minor Enchantments & 2–4 XP \\
Major Enchantments & 6–8 XP \\
Prestige Enchantments & 10+ XP \\
\bottomrule
\end{longtable}

\noindent \textbf{Limitations:}
\begin{itemize}
    \item Max enchantments = Spirit attribute.
    \item No stacking identical benefits.
    \item Maintenance affects functionality.
\end{itemize}

\section{Asset and Follower Management}\label{sec:asset_management}
\index{Assets}\index{Followers}\index{Resources!Assets and Followers}

\subsection{Followers (On‑Screen Resources)}\label{subsec:followers}
\index{Followers!On-Screen}

\begin{description}
    \item[\indexterm{Cost}] Cap\(^{2}\) XP to acquire.\index{Followers!Cost}\index{XP!Followers}
    \item[\indexterm{Assist Dice}] When applicable, the follower adds help dice equal to \(\min(\text{Cap},\ \text{helper's relevant Skill})\), capped at \textbf{+3} dice total from all sources. \emph{Exception:} \textit{Exceptional Coordination} Talent allows one follower to provide \textbf{+4} assist dice.\index{Assists}\index{Talents!Exceptional Coordination}
    \item[\indexterm{Capability}] Ranges from 1–5 (5 is exceptional).\index{Followers!Capability}
    \item[\indexterm{Upkeep}] Each Downtime period, pay XP equal to Cap \emph{or} spend a Scene tending the relationship.\index{Downtime!Followers}\index{XP!Upkeep}
    \item[\indexterm{Risk}] If the GM spends 2+ SB on an action you take with assistance, they may endanger, injure, or separate the follower instead of you if fictionally appropriate.\index{Story Beats!Follower risk}
    \item[\indexterm{Off‑Screen Capability}] Once per downtime, a Cap 5 follower can solve one significant problem but generates 1 SB for the party; the GM must describe how their action creates consequences.\index{Followers!Off-Screen}
\end{description}

\subsection{Upkeep (Expanded)}\label{sec:upkeep_expanded}
\index{Upkeep}\index{Downtime}\index{Followers}\index{Assets}\index{Wary}\index{Seized}\index{Neglected}\index{Compromised}

Upkeep is the story of attention. Followers and assets thrive when seen and sour when ignored.

\subsubsection{Design Intent}\label{subsec:upkeep_intent}
\index{Design Philosophy}
\begin{itemize}
    \item \textbf{Choice with Teeth.} Players choose \emph{time} or \emph{XP}. Either way, the fiction moves: quick delegation vs. hands‑on scenes.
    \item \textbf{Visible Drift.} Missing upkeep degrades things predictably: \textsc{Wary} $\rightarrow$ \textsc{Seized}, \textsc{Neglected} $\rightarrow$ \textsc{Compromised}.
    \item \textbf{Story First.} Every payment should say something about the relationship or the tool.
\end{itemize}

\subsubsection{Asset Upkeep}\label{subsec:upkeep_core}
\textbf{Frequency.} Once per Downtime period.

\begin{description}
    \item[\textbf{Option A — Efficient (Higher XP, Less Time)}] Pay Upkeep XP \(= \max\bigl(1, \frac{\text{XP Acquisition Cost}}{3}\bigr)\). Minimal effort in‑fiction (a retainer handles it; you check in by letter).
    \item[\textbf{Option B — Intensive (Lower XP, More Time)}] Pay 1 XP. Spend a \emph{dedicated Downtime action} personally tending the follower / asset.
\end{description}

\textbf{Failure.} If unpaid this Downtime:
\begin{itemize}
    \item \emph{Follower:} becomes \textbf{Wary} (or \textbf{Seized} if already Wary).
    \item \emph{Asset:} becomes \textbf{Neglected} (or \textbf{Compromised} if already Neglected).
\end{itemize}

\subsubsection{What the Conditions Look Like}\label{subsec:upkeep_conditions}
\begin{description}
    \item[\textsc{Wary} (Follower)] They hesitate, ask for guarantees, or interpret orders narrowly. \emph{Mechanical nudge:} first social / test involving them is \(-1\) position \emph{or} costs 1 extra Stress.
    \item[\textsc{Seized} (Follower)] Someone else sets terms (rival, guild, family) or they withdraw until appeased. \emph{Nudge:} cannot assist this Downtime; to re‑engage, pay 1 Boon \emph{or} clear via a narrative scene that addresses their grievance.
    \item[\textsc{Neglected} (Asset)] It underperforms; parts go missing; paperwork piles. \emph{Nudge:} \(-1\) effect on its next use \emph{or} an \emph{Alarmed Attention} clock +1/2.
    \item[\textsc{Compromised} (Asset)] Faults propagate; someone has leverage; the tool is noisy or unsafe. \emph{Nudge:} first use this session automatically creates 1 \textbf{SB} \emph{or} requires a costly fix scene.
\end{description}

\subsubsection{Examples by Type}\label{subsec:upkeep_examples}

\paragraph{Followers.}
\begin{longtable}{p{2.5cm} p{5cm} p{5cm}}
\caption{Follower Upkeep Examples}\label{tab:follower_upkeep}\\
\toprule
\textbf{Type} & \textbf{Efficient Upkeep (pay XP, quick)} & \textbf{Intensive Upkeep (1 XP + action)} \\
\midrule
Scribe & Send coin and a sealed brief with instructions. & Host a lesson; co‑author a pamphlet that sharpens their craft. \\
Scout & Dispatch a courier with a new map and stipend. & Walk the route together; overhaul signals and caches. \\
Acolyte (Patron‑bound) & Arrange a blessing via intermediaries. & Lead them in a minor rite; narrate the teaching moment. \\
Mercenary & Wire hazard pay with a short commendation. & Drill formation; settle a dispute; toast the unit. \\
Informant & Drop a dead‑letter and payment. & Meet face‑to‑face; share protection protocol. \\
\bottomrule
\end{longtable}

\paragraph{Assets.}
\begin{longtable}{p{2.5cm} p{5cm} p{5cm}}
\caption{Asset Upkeep Examples}\label{tab:asset_upkeep}\\
\toprule
\textbf{Type} & \textbf{Efficient Upkeep (pay XP, quick)} & \textbf{Intensive Upkeep (1 XP + action)} \\
\midrule
Workshop & Pay the foreman; authorise routine parts. & Personally recalibrate; craft a sample piece. \\
Safehouse & Hire a caretaker; basic supplies delivered. & Repair locks/windows; rewrite the cover story with neighbours. \\
Library & Fund copying; buy indices. & Catalogue a shelf; cross‑ref a case; bind a damaged tome. \\
Boat & Pay moorage and minimal maintenance. & Scrape hull; patch sails; sail a proving run. \\
Sigil Network (Invoker) & Replace two worn plates. & Etch a new master glyph; clear 1 wear mid‑scene next session. \\
\bottomrule
\end{longtable}

\subsubsection{Edge Cases \& Rulings}\label{subsec:upkeep_edge}
\begin{itemize}
    \item \textbf{Multiple Resources.} Each follower / asset checks upkeep separately. A single Intensive action can cover a \emph{cohesive group} if fiction supports it (e.g., one drill for a squad).
    \item \textbf{Remote Care.} If you narrate convincing remote oversight (sending a trusted lieutenant), count as Efficient.
    \item \textbf{Patron‑Tinted Upkeep.} Tie upkeep scenes to a Patron’s theme for small boons (advantage on the next related roll) without changing costs.
    \item \textbf{Stacking Misses.} Missing upkeep across \emph{two} consecutive Downtime periods moves \textsc{Wary} $\rightarrow$ \textsc{Seized} or \textsc{Neglected} $\rightarrow$ \textsc{Compromised}; a third consecutive miss risks loss (follower quits; asset condemned) at GM discretion.
\end{itemize}

\subsubsection{Optional Modules}\label{subsec:upkeep_modules}
\index{Ledger Credits}\index{Union Rules}\index{Patron Favor Swap}

\begin{description}
    \item[\textbf{Ledger Credits}] Track fractional prep: three Efficient payments can be banked to waive one Intensive action later.
    \item[\textbf{Union Rules}] Some factions demand Intensive upkeep; Efficient counts as half (two Efficient = one Intensive).
    \item[\textbf{Patron Favor Swap}] Spend a \textbf{Boon} to treat missed upkeep as paid for one follower / asset if the scene honours the Patron.
\end{description}

\subsubsection{GM Guidance}\label{subsec:upkeep_gm}
\begin{itemize}
    \item \textbf{Making it a Scene.} Intensive upkeep should reveal something (new contact, flaw, rumour).
    \item \textbf{Foreshadowing Degradation.} Before applying \textsc{Seized} / \textsc{Compromised}, show warning signs the players can act on.
    \item \textbf{Naming the Toll.} If a \textbf{Patron’s Largess} (wrath) applies, state the narrative reason and the extra toll (e.g., an added SB, a temporary lockout, or a vow demanded).
\end{itemize}

\begin{tcolorbox}[title={Upkeep Quick Reference},colback=gray!5,colframe=black]
\textbf{Once per Downtime.} Pay Efficient (XP = \(\max(1, \frac{\text{acq}}{3})\)) or Intensive (1 XP + action). \\
\textbf{Missed:} Follower \(\to\) \textsc{Wary} \(\to\) \textsc{Seized}; Asset \(\to\) \textsc{Neglected} \(\to\) \textsc{Compromised}. \\
Tie scenes to your Patron for flavour, not discounts.
\end{tcolorbox}

\subsection{Follower Assist Rules}\label{subsec:follower_assist}
\index{Followers!Assist}\index{Assists!Rules}

\begin{itemize}
    \item Assist dice come from the helper’s capabilities, not the leader’s.\index{Assists!Source}
    \item Total Assist on any roll (from any sources) is hard‑capped at \textbf{+3}. \emph{Exception:} \textit{Exceptional Coordination} may allow \textbf{+4} from a single follower.\index{Assists!Maximum}
    \item Only one follower may assist a given action at a time.\index{Followers!Slot limit}
    \item Followers cannot assist actions beyond their narrative scope or capabilities.
\end{itemize}

\subsection{Loyalty \& Bonds (Optional Rules)}\label{subsec:follower_loyalty}
\index{Followers!Loyalty}\index{Bonds}

\begin{itemize}
    \item Track a simple Loyalty tag per follower: \emph{Wary / Steady / Devoted}.\index{Followers!Loyalty tags}
    \item \emph{Devoted} followers can once per arc convert one GM Complication targeting them into a lesser setback.\index{Complications}
    \item \emph{Wary} followers cost +1 XP to maintain during Downtime.\index{XP!Upkeep}
    \item Loyalty can change based on how the PC treats the follower and shared experiences.
\end{itemize}

\subsection{Off‑Screen Assets}\label{subsec:off_screen_assets}
\index{Assets!Off-Screen}

\begin{longtable}{p{4cm} p{9cm}}
\caption{Asset Types and Costs}\label{tab:asset_types}\\
\toprule
\textbf{Asset Type} & \textbf{Cost} \\
\midrule
Minor Asset & 4 XP (safehouse, small charter, local business) \\
Standard Asset & 8 XP (noble title, guild section, spy ring, significant property) \\
Major Asset & 12 XP (city licence, regional network, major institution influence) \\
Artifact Asset & 16+ XP (unique items or positions with campaign‑level significance) \\
\bottomrule
\end{longtable}

\subsubsection{Asset Activations}\label{subsec:asset_activations}
\index{Assets!Activations}\index{Boons}

\begin{itemize}
    \item \textbf{Off‑Screen Activation:} At campaign start or during Downtime, activate an off‑screen asset by spending \textbf{1 Boon} or \textbf{2 XP}.\index{Assets!Activation cost}\index{XP!Assets}
    \item \textbf{Off‑Screen Effects:} Use each Asset’s listed off‑screen effect once per session for free.\index{Assets!Off-Screen effects}
    \item \textbf{On‑Screen Activation:} To reshape the current scene, spend \textbf{1 Boon}.\index{Assets!On-Screen effects}
    \item \textbf{Plausibility Test:} The Asset must have appropriate scope and reach for the intended effect.\index{Assets!Plausibility}
\end{itemize}

\subsubsection{Asset Condition Tracks}\label{subsec:asset_conditions}
\index{Assets!Condition tracks}

\begin{description}
    \item[\indexterm{Maintained}] Full capability; no penalties. The asset is in good standing and fully functional.\index{Assets!Maintained}
    \item[\indexterm{Neglected}] \(-1\) die when used (assist or leverage). Narratively: slower response, sullen staff, short‑staffed operations.\index{Assets!Neglected}
    \item[\indexterm{Compromised}] Unavailable for use. Narratively: captured, burned, seized, defected, or otherwise incapacitated.\index{Assets!Compromised}
\end{description}

\subsubsection{Maintenance and Repair}\label{subsec:asset_maintenance}
\index{Assets!Maintenance}\index{Downtime}

\begin{itemize}
    \item After a story arc or 2–3 sessions of heavy use, the GM may flag the resource as \emph{at risk}.\index{Assets!Risk flag}
    \item To keep an asset \emph{Maintained}, the player must either:
        \begin{itemize}
            \item Spend Downtime repairing / servicing it, or\index{Assets!Downtime repair}
            \item Pay XP immediately (representing replacement parts, masterwork upkeep).\index{Assets!XP repair}
        \end{itemize}
    \item If neglected, the asset becomes \emph{Neglected} (\(-1\) die) and may eventually become permanently lost.\index{Assets!Breakage}
    \item Superior and Artifact assets do not require normal upkeep; if \emph{Compromised} through complications, only narrative quests can repair them.\index{Artifacts}\index{Quests}
\end{itemize}

\section{Bond‑Driven Resource Generation}\label{sec:bond_resources}
\index{Bonds}\index{Boons}\index{Resources!Bond-Driven Generation}

When a player takes a significant action to aid an ally with whom they share a bond, and explicitly references that bond in an \emph{Intricate} description, they may mark that bond to gain \textbf{1 Boon} \emph{after} the action resolves.\index{Description Ladder!Intricate action}

\subsection{Requirements}\label{subsec:bond_requirements}
\index{Bonds!Requirements}

\begin{itemize}
    \item \textbf{Mutual Bond:} Player shares a defined bond with the ally they are aiding.\index{Bonds!Mutual requirement}
    \item \textbf{Intricate Description:} Player describes how the bond motivates their action using rich, multi‑sensory details.\index{Description Ladder!Intricate action}
    \item \textbf{Significant Aid:} Meaningful assistance beyond basic dice bonuses.\index{Teamwork}\index{Assists}
    \item \textbf{Fiction First:} The bond genuinely drives the choice to help, not added retroactively.\index{Fiction-first}
\end{itemize}

\subsection{Examples}\label{subsec:bond_examples}
\index{Bonds!Examples}

\textbf{Valid Examples:}
\begin{itemize}
    \item “Remembering how they saved me from the falling rubble in Aeler, I throw myself in front of the crossbow bolt meant for them!”
    \item “Thinking of our shared vow to protect the innocent, I use my last healing potion to stabilise them instead of saving it for myself.”
    \item “Drawing on our years fighting side‑by‑side in the Border Wars, I rally the other mercenaries to keep fighting alongside them when morale fails.”
\end{itemize}

\subsection{Limitations}\label{subsec:bond_limitations}
\index{Bonds!Limits}

\begin{itemize}
    \item Once per bond per session.\index{Bonds!Session limit}
    \item Must involve meaningful sacrifice or risk.\index{Risk}
    \item GM approval required for what constitutes “significant action”.\index{GM discretion}
    \item Cannot be used for basic assistance rolls or minor favours.\index{Description Ladder!Basic action}
    \item The Boon is awarded \emph{after} the action resolves, not before.\index{Boons!Timing}
\end{itemize}

\section{Over‑Stack Rule}\label{sec:over_stack}
\index{Combat!Over-Stack}\index{Advantages}

The Over‑Stack rule prevents excessive accumulation of advantages from trivial sources while rewarding meaningful preparation.

\subsection{Structural Advantages}\label{subsec:structural_advantages}
\index{Combat!Structural advantages}

Structural advantages include:
\begin{itemize}
    \item Active buffs or beneficial tags affecting the party.\index{Tags}
    \item Favourable venue or environmental factors.\index{Environment}
    \item Unused \emph{Follower Initiative} for the scene.\index{Followers!Initiative}
    \item On‑screen Asset activation providing immediate benefits.\index{Assets!On-Screen effects}
    \item Opponent disadvantaged by fiction (surprised, trapped, etc.).\index{Position!Surprise}
    \item Ritual preparation that applies to the current situation.\index{Rituals!Preparation}
\end{itemize}

\subsection{Over‑Stack Trigger}\label{subsec:overstack_trigger}
\index{Combat!Over-Stack trigger}\index{Story Beats}

\begin{itemize}
    \item If the party \textbf{enters} a scene with \(\geq 3\) structural advantages, a GM \textit{may} apply Over‑Stack once for that scene.
    \item The GM chooses either:
        \begin{itemize}
            \item Start one named obstacle or challenge at \textbf{+1 DV}, or\index{Difficulty Value}
            \item The GM banks \textbf{+1 SB} for the first Deck Twist in the scene.\index{Deck of Consequences!Twists}
        \end{itemize}
    \item This represents narrative pushback when characters have overwhelming advantages.
    \item Over‑Stack applies only once per scene, regardless of how many advantages accumulate.
\end{itemize}

\noindent \textit{Note:} Over‑Stack applies to advantages held at scene framing, not those generated through play. If the party enters with overwhelming force, the world pushes back once. If they create overwhelming force through creative improvisation during the scene, that is the intended reward cycle—no Over‑Stack applies.

\section{Currency-Free Economy (Compressed Reference)}
\label{sec:currency-free-ref}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!70!black,title=Quick Reference]
\emph{No coin. Incidental expenses are hand-waved. Significant costs use XP, barter, Strings, or Obligation (Debt Clocks).}
\end{tcolorbox}

\subsection{What Costs What}
\label{ref:costs-table}

\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Item / Benefit} & \textbf{XP Cost} & \textbf{Alternative Payment} \\
\midrule
Minor Asset & 4 XP & Minor String or Debt Clock [4] \\
Standard Asset & 8 XP & Standard String or Debt Clock [6] \\
Major Asset & 12 XP & Major String or Debt Clock [8] \\
Follower (Cap $n$) & $n^2$ XP & Debt Clock [$n+2$] or equivalent String \\
Single-use magical item & 2–4 XP & Barter (item of equal rarity) \\
Permanent magical item (Tier I) & 6–10 XP & Quest + String \\
Clear Debt Clock segment & 2 XP & One service scene \\
Permanent String (earned) & — (reward) & — \\
Form a Trust (party pool) & 10 XP & — \\
\bottomrule
\end{tabular}
\end{center}

\subsection{String Values}
\label{ref:string-values}

\begin{center}
\small
\begin{tabular}{ll}
\toprule
\textbf{String Tier} & \textbf{Purchasing Power} \\
\midrule
Minor & 1 XP discount, or one free use of a Minor Asset effect \\
Standard & Minor Asset (4 XP) or clear 2 Debt Clock segments \\
Major & Standard Asset (8 XP) or Cap 3 Follower \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Debt Clock Equivalences}
\label{ref:debt-clock}

\begin{center}
\small
\begin{tabular}{ll}
\toprule
\textbf{Debt Clock Size} & \textbf{Equivalent Value} \\
\midrule
[4] & Minor Asset or one-time service \\
[6] & Standard Asset or Cap 3 Follower \\
[8] & Major Asset or Cap 5 Follower \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Trust Investments (from Party Pool)}
\label{ref:trust-invest}

\begin{center}
\small
\begin{tabular}{ll}
\toprule
\textbf{XP} & \textbf{Effect} \\
\midrule
4 & Increase Trust Debt Clock capacity by 2 \\
6 & Add Minor Asset (Trust-owned) \\
8 & Upgrade Asset (Minor → Standard) \\
10 & Add Cap 3 Follower (Trust lieutenant) \\
12 & Gain permanent Standard String (faction owes Trust) \\
15 & Trust becomes Major Terrestrial Patron (can grant Rites) \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Hand-Waved (Free)}
\label{ref:free}

Meals, inn stays, basic tools, riding horse, routine transport, minor repairs, small bribes.

\subsection{GM Adjudication Quick Mat}
\label{ref:gm-mat}

\begin{center}
\small
\begin{tabular}{p{3cm}p{10cm}}
\toprule
\textbf{Situation} & \textbf{Ruling} \\
\midrule
Player asks for small, reasonable item & Yes, free. \\
Player wants a permanent asset & XP, String, or Debt Clock. \\
Player offers barter & Roll Presence+Sway (DV 2–5); on success, trade happens. \\
Player spends a String & No roll unless hostile/extreme; String is consumed. \\
Debt Clock fills & Creditor makes a demand (mission, sacrifice, public oath). \\
Player hoards XP & Introduce problems that need Assets/Trusts, not personal power. \\
Player farms Boons for XP & Anti-fishing: no Boons from safe/repetitive actions. \\
Player founds a Trust & Require 10 XP from pool; run introductory mission to clear founding Debt Clock. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{SB Menu – Currency Complications}
\label{ref:sb-currency}

\begin{center}
\small
\begin{tabular}{ll}
\toprule
\textbf{SB Spent} & \textbf{Complication} \\
\midrule
1 & Barter item damaged; NPC demands extra favor. \\
2 & Debt Clock ticks unexpectedly; String found counterfeit. \\
3 & Trust Asset threatened; major barter deal collapses. \\
4+ & Trust schism; Debt Clock called in early at worst moment. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Currency-Free Economy – Core Principles}
\label{ref:principles}

\begin{itemize}
  \item \textbf{Hand-wave trivial.} Say yes to small, sensible requests.
  \item \textbf{Significant gains cost XP, Strings, or Debt.}
  \item \textbf{Debt is story.} Call in Debt Clocks at dramatic moments.
  \item \textbf{Trusts are player-driven factions.} Use them to generate missions and rivals.
  \item \textbf{Spent XP is remembered.} Make assets and followers matter in the fiction.
\end{itemize}