\chapter{Character Advancement}
\label{chap:character-advancement}
\index{Advancement}\index{Experience Points (XP)}\index{Character Creation}

\section{Introduction to Advancement}
\label{sec:advancement-intro}
\index{Advancement!Overview}

Character advancement in \textbf{Fate's Edge} reflects meaningful growth in capability and standing. XP is awarded for table-facing accomplishments, hard choices, and dramatic friction that defines your story. Advancement should always reinforce narrative identity: what drives, obligations, and affinities are shaping the character? \index{XP!Awards}\index{Narrative identity}

\section{Starting Character Builds}
\label{sec:starting-builds}
\index{Character Creation!Starting Builds}

\subsection{Standard Starting XP}
\label{subsec:standard-starting}
\index{Character Creation!30 XP}

Players begin with \textbf{30 XP} to allocate during initial character creation. This represents a balanced baseline for competent starting characters.

\subsection{Enhanced Starting Options}
\label{subsec:enhanced-starting}
\index{Character Creation!Enhanced Builds}\index{Bonds}\index{Complications}

Players may exceed the standard 30 XP build through narrative engagement:
\begin{itemize}
\item \textbf{Bonds:} Up to two player-defined mutual bonds may be taken for +2 XP total (+1 XP each).\index{Bonds!Starting Bonus}
\item \textbf{Complications:} Up to two initial complications may be accepted for +4 XP total (+2 XP each).\index{Complications!Starting Bonus}
\end{itemize}

This allows for a maximum starting build of \textbf{34 XP}. Players are encouraged to aim for 30 XP and use bonds/complications to enhance characterization rather than pure mechanical optimization. \index{Character Creation!34 XP cap}

\subsection{Initial Complications Reminder}
\label{subsec:initial-complications}
\index{Complications!Initial}

For each initial complication taken at character creation:
\begin{itemize}
\item Start each scene with +1 banked SB per character with initial complications.\index{Story Beats (SB)!Banked}
\item These complications remain until cleared through play or character development.\index{Complications!Clearing}
\end{itemize}

\section{Player Archetypes}
\label{sec:player-archetypes}
\index{Archetypes}\index{Build styles}

Not every group plays the same way. These archetypes describe how players typically spend XP and seek spotlight.

\subsection{The Solo}
\label{subsec:solo-archetype}
\index{Archetypes!Solo}

\begin{description}
\item[\textbf{Definition}] Invests XP primarily in Attributes and Skills. Minimal followers, minimal holdings. All power is on the character sheet.\index{Attributes}\index{Skills}
\item[\textbf{Typical XP Spread}] 70--90\% Self; 0--10\% On-screen help; 0--20\% Off-screen assets.\index{XP!Spreads}
\item[\textbf{Strengths}] Consistent scene impact; few moving parts; resilient to follower loss.\index{Followers!Loss resilience}
\item[\textbf{Risks}] Limited fiction reach between sessions; can stall when problems demand logistics or networks.\index{Logistics}\index{Networks}
\item[\textbf{Starting Focus}] Attributes 2--3, Skills 1--2, minimal assets.\index{Assets!Minimal}
\end{description}

\subsection{The Mixed Player}
\label{subsec:mixed-archetype}
\index{Archetypes!Mixed}

\begin{description}
\item[\textbf{Definition}] Splits XP between self-growth and one or two meaningful assets (a small follower or a reliable holding).\index{Assets}\index{Followers}
\item[\textbf{Typical XP Spread}] 50--65\% Self; 15--25\% On-screen help; 15--25\% Off-screen assets.\index{XP!Spreads}
\item[\textbf{Strengths}] Versatile: credible in scenes and has a lever for special problems.\index{Versatility}
\item[\textbf{Risks}] Upkeep pressure; helper can be targeted when the GM spends 2+ Story Beats.\index{Upkeep}\index{Story Beats (SB)!GM spends}
\item[\textbf{Starting Focus}] Balanced approach with one minor asset or low-cap follower.\index{Assets!Minor}\index{Followers!Cap}
\end{description}

\subsection{The Mastermind}
\label{subsec:mastermind-archetype}
\index{Archetypes!Mastermind}

\begin{description}
\item[\textbf{Definition}] Prioritizes followers/cadres/familiars and off-screen networks. The character sheet is the hub of a larger apparatus.\index{Networks}\index{Cadres}\index{Familiars}
\item[\textbf{Typical XP Spread}] 25--40\% Self; 35--55\% On-screen help; 20--40\% Off-screen assets.\index{XP!Spreads}
\item[\textbf{Strengths}] Scene control via assistance; strategic reach between sessions; strong heist/social-planning play.\index{Teamwork}\index{Planning}
\item[\textbf{Risks}] Dependency on assist lanes; followers can be endangered on 2+ SB spends; upkeep pressure.\index{Assists}\index{Followers!Risk}
\item[\textbf{Starting Focus}] Significant investment in followers or major assets from the beginning.\index{Assets!Major}
\end{description}

\section{XP Costs and Advancement}
\label{sec:xp-costs}
\index{XP!Costs}\index{Advancement!Costs}

\subsection{Core Character Improvements}
\label{subsec:core-improvements}
\index{Attributes!Costs}\index{Skills!Costs}\index{Followers!Costs}

\begin{description}
\item[\textbf{Attributes}] Cost = new rating $\times$ 3 XP. Requires downtime equal to new rating in days.\index{Downtime!Advancement}
\item[\textbf{Skills}] Cost = new level $\times$ 2 XP. Requires downtime equal to new level in days.
\item[\textbf{On-Screen Followers}] Cost = Cap$^{2}$ XP. Requires 1--3 days downtime to recruit and brief.
\end{description}

\subsection{Off-Screen Assets}
\label{subsec:off-screen-assets}
\index{Assets!Costs}

\begin{description}
\item[\textbf{Minor Asset}] 4 XP; requires 1 day downtime to establish.\index{Assets!Minor}
\item[\textbf{Standard Asset}] 8 XP; requires 1 week downtime to establish.\index{Assets!Standard}
\item[\textbf{Major Asset}] 12 XP; requires 1 month downtime to establish.\index{Assets!Major}
\item[\textbf{Artifact Asset}] 16+ XP; requires significant campaign effort to acquire.\index{Artifacts}
\end{description}

\subsection{Talent Costs}
\label{subsec:talent-costs}
\index{Talents!Costs}

\begin{description}
\item[\textbf{Minor Talents}] 2 XP --- Small narrative tricks, situational bonuses.\index{Talents!Minor}
\item[\textbf{Major Talents}] 4 XP --- Strong abilities, permanent +1 effects in niches.\index{Talents!Major}
\item[\textbf{Prestige Abilities}] 6+ XP --- Campaign-defining effects, rare and powerful.\index{Prestige Abilities}
\end{description}

\section{Tiers of Reputation}
\label{sec:reputation-tiers}
\index{Reputation Tiers}

Though there are no character levels, total XP spent creates soft ``tiers'' that shape how the world responds to characters:\index{XP!Totals}

\begin{description}
\item[\textbf{Tier I -- Rookie}] (0--40 XP): Local reputation; prestige abilities locked.\index{Reputation Tiers!Rookie}
\item[\textbf{Tier II -- Seasoned}] (41--90 XP): Regional notice; prestige abilities may unlock.\index{Reputation Tiers!Seasoned}
\item[\textbf{Tier III -- Veteran}] (91--150 XP): National influence; second follower slot suggested.\index{Reputation Tiers!Veteran}
\item[\textbf{Tier IV -- Paragon}] (151--220 XP): Movers and shakers; rivals emerge to challenge.\index{Reputation Tiers!Paragon}
\item[\textbf{Tier V -- Mythic}] (221+ XP): Legendary status; kingdoms and cults respond directly.\index{Reputation Tiers!Mythic}
\end{description}

\section{Prestige Abilities and Talents}
\label{sec:prestige-abilities}
\index{Talents}\index{Prestige Abilities}

Prestige abilities are narrative milestones unlocked through mastery or story events. They are priced at 6+ XP and include campaign-shaping effects. \index{Milestones}

\subsection{General Talents}
\label{subsec:general-talents}
\index{Talents!General}

\begin{description}
\item[\textbf{Battle Instincts}] (Cost: 6 XP): Once per scene, re-roll a failed defense roll.\index{Battle Instincts}
\item[\textbf{Silver Tongue}] (Cost: 4 XP): Gain +1 die when persuading or deceiving through speech.\index{Silver Tongue}
\item[\textbf{Iron Stomach}] (Cost: 6 XP): Immune to mundane poisons and spoiled food; halve complications from toxic sources.\index{Iron Stomach}
\item[\textbf{Exceptional Coordination}] (Cost: 8 XP): One follower can provide +4 assist dice (exception to the +3 cap).\index{Exceptional Coordination}\index{Assists!Cap}
\end{description}

\subsection*{Subtle Casting (Major Talent --- 8 XP)}
\index{Talents!Subtle Casting}

\textbf{Requirements:} Lore 3+, Performance 2+ \emph{or} Runekeeper with Codex

\textbf{Effect:} Make a \textbf{Performance + Lore} roll to quietly cast a spell, invoke a Rite, or sing a Cantos against DV (Tier). 
If successful, the casting does not generate \index{Story Beats} on the \emph{Channel} or initial roll. 
This talent allows the caster to veil magic in story, song, or symbol rather than force.

\textbf{Limitations:}
\begin{itemize}
  \item Cannot be used for \emph{Great} or \emph{Extreme} Tier effects.
  \item The \emph{Weave} phase (if applicable) still generates normal SB.
  \item Obvious magical manifestations still occur (glowing sigils, strange sounds, sudden winds, etc.).
\end{itemize}

\begin{quote}
\emph{"True subtlety is not silence, but harmony --- when even the wind believes it sang the song."}
\end{quote}


\paragraph{Backstab (Major Talent, 8 XP)} 
\textbf{Req:} Stealth~2+, Melee~2+, Light weapon.  

\textbf{Effect:} When you attack an \textbf{Unaware} or \textbf{Engaged} foe from \textbf{Stealth}, deal \emph{+1 Harm} and ignore \emph{1 point of their Armor}.  

\textbf{Definitions:}  
\begin{itemize}
  \item \textbf{Unaware:} The target is not aware of your presence or hostile intent. This typically requires being \emph{Hidden} or having succeeded on a \emph{Stealth} test.  
  \item \textbf{Engaged:} The target is currently taking an \emph{Attack} action against another character, or casting a spell/ritual that specifically targets another character.  
  \begin{itemize}
    \item In miniatures/tactical play: the target is in melee range (\emph{Close}) with another PC/NPC and actively fighting them.  
  \end{itemize}
\end{itemize}

\textbf{Limit:} Once per scene. To use again, you must first \emph{re-enter Stealth} (DV by narrative) and mark 1 \emph{Fatigue} (e.g., via \emph{Shadow Dance}).  

\textbf{On a Miss:} You are \emph{Exposed} --- drop to \emph{Desperate Position} or mark 1 \emph{Harm}.    

\paragraph{Shadow Dance (Synergy Talent, 10 XP)}  
\textbf{Req:} Backstab, Stealth~3+, Mobility~2+.  

\textbf{Effect:} After a successful \textbf{Backstab}, you may immediately test \textbf{Stealth} vs. DV (Tier).  
\begin{itemize}
  \item On success: You \emph{re-enter Stealth} and may either \textbf{clear 1 Fatigue} or \textbf{improve Position +1}.  
  \item On failure: You remain \emph{Exposed} and must mark 1 \emph{Fatigue}.  
\end{itemize}  

\textbf{Limit:} May only chain once per scene.   

\paragraph{Deathblow (Capstone Talent, 12 XP)}  
\textbf{Req:} Shadow Dance, Stealth~4+, Melee/Ranged~3+.  

\textbf{Effect:} When you strike from \textbf{Dominant Position} or after re-entering \textbf{Stealth} via \emph{Shadow Dance}, you may declare a \textbf{Deathblow}.  
\begin{itemize}
  \item On a hit: Deal \emph{triple Harm}. If the attack incapacitates the target, you may immediately attempt a free \textbf{Stealth} test (DV by narrative) to vanish.  
  \item On a miss: You are \emph{Exposed} — drop to Desperate Position and mark 1 \emph{Harm}.  
\end{itemize}  

\textbf{Limit:} Once per scene. You may mark 1 \emph{Fatigue} to attempt a second time.  

\subsection{Cultural and Ancestral Talents}
\label{subsec:cultural-talents}
\index{Talents!Cultural}

\subsubsection{Humans}
\label{subsubsec:human-talents}
\index{Humans!Talents}

\begin{description}
\item[\textbf{Versatile}] (3 XP): Once per session, swap one Skill die for another you lack.\index{Versatile (Human)}
\item[\textbf{Guild Ties}] (7 XP; Requires Presence 2+): Call in favors from a guild or order once per session.\index{Guild Ties (Human)}\index{Presence}
\item[\textbf{Banner-Bearer}] (15 XP; Requires Presence 4+, Resolve 3): Inspire an allied unit; once per battle, erase 2 complications for the group.\index{Banner-Bearer (Human)}\index{Resolve}
\end{description}

\subsubsection{Dwarves (Aeler)}
\label{subsubsec:dwarf-talents}
\index{Dwarves!Talents}\index{Aeler}

\begin{description}
\item[\textbf{Stone-Sense}] (5 XP; Dwarf only): Detect flaws in stone; +1 die to underground \emph{Lore} or \emph{Survival} checks involving stonework.\index{Stone-Sense (Dwarf)}\index{Lore}\index{Survival}
\item[\textbf{Spirit Shield}] (8 XP; Requires Spirit 3+): Commune with ancestors; once per session, block 1 complication for an ally.\index{Spirit Shield (Dwarf)}\index{Spirit}
\item[\textbf{Forge-Patriarch}] (18 XP; Requires Body 4+, Lore 3): Found a forge-citadel; gain loyal smiths and engineers (Cap 5 followers).\index{Forge-Patriarch (Dwarf)}\index{Followers!Cap 5}\index{Body}
\end{description}

\subsubsection{Wood Elves}
\label{subsubsec:wood-elf-talents}
\index{Wood Elves!Talents}

\begin{description}
\item[\textbf{Backlash Soothing}] (5 XP; Wood Elf only): Once per session, cancel 1 Backlash die in natural terrain.\index{Backlash Soothing (Wood Elf)}\index{Backlash}
\item[\textbf{Ranger's Step}] (9 XP; Requires Wits 3+, Stealth 2): Move unseen in forests; treat terrain penalties as one step lower.\index{Ranger's Step (Wood Elf)}\index{Wits}\index{Stealth}
\item[\textbf{Wild Speaker}] (18 XP; Requires Spirit 5+, Survival 3): Speak to beasts and trees; once per session, command local wildlife as allies.\index{Wild Speaker (Wood Elf)}\index{Survival}
\end{description}

\subsubsection{High Elves}
\label{subsubsec:high-elf-talents}
\index{High Elves!Talents}

\begin{description}
\item[\textbf{Lorekeeper}] (4 XP; High Elf only): Recall obscure history or magic without rolling once per session.\index{Lorekeeper (High Elf)}
\item[\textbf{Weave Anchor}] (8 XP; Requires Wits 3+, Arcana 3): Reduce magical Backlash by 1 when casting Intricate spells.\index{Weave Anchor (High Elf)}\index{Arcana}
\item[\textbf{Echo-Walker's Step}] (20 XP; High Elf only; Requires Wits 5, Arcana 4): Step briefly into the Ways Between; once per arc, turn a complication into a boon.\index{Echo-Walker's Step (High Elf)}
\end{description}

\subsubsection{Ykrul}
\label{subsubsec:ykrul-talents}
\index{Ykrul!Talents}

\begin{description}
\item[\textbf{Blood Frenzy}] (4 XP; Requires Body 2+): When reduced to half health, gain +1 die on melee rolls.\index{Blood Frenzy (Ykrul)}
\item[\textbf{Blood Memory}] (7 XP; Requires Body 3+): After battle, gain 1 temporary Skill die reflecting a foe's tactics in the next scene.\index{Blood Memory (Ykrul)}
\item[\textbf{Warglord}] (18 XP; Requires Body 5, Sway 3): Rally scattered warbands; once per campaign, unify tribes under one banner.\index{Warglord (Ykrul)}\index{Sway}
\end{description}

\subsection{High-Tier Prestige Abilities}
\label{subsec:high-tier-prestige}
\index{Prestige Abilities!High tier}

\begin{description}
\item[\textbf{Echo-Walker's Step}] (High Elf; Cost: 20 XP; Requires Wits 5, Arcana 4): Once per arc, observe a perfect echo of a past event at your location. GM immediately banks +2 SB; scenes touching that memory carry an omen. Grants DV $-1$ on one action that uses the revealed truth.\index{Echo-Walker's Step}\index{Omens}\index{Difficulty Value (DV)}
\item[\textbf{Warglord}] (Ykrul; Cost: 18 XP; Requires Body 5, Sway 3): Once per campaign, unify scattered warbands into a single host for a season. Start a Logistics clock and a Grudge clock; either one filling fractures the host.\index{Warglord}\index{Clocks}
\item[\textbf{Spirit-Shield}] (Aeler; Cost: 15 XP; Requires Spirit 4, Resolve 3): Once per session, erase up to 3 SB from an ally's current roll; you immediately mark Fatigue +1 and the GM banks +1 SB as backlash.\index{Spirit-Shield}\index{Resolve}\index{Fatigue}\index{Story Beats (SB)}
\end{description}

\section{Starting Build Examples}
\label{sec:starting-examples}
\index{Character Creation!Examples}\index{Build examples}

\subsection{Solo Build (30--32 XP)}
\label{subsec:solo-example}
\index{Archetypes!Solo!Example}

\begin{itemize}
\item Body 3 (9 XP), Wits 2 (6 XP), Spirit 2 (6 XP).\index{Body}\index{Wits}\index{Spirit}
\item Skills: Combat 2 (4 XP), Stealth 1 (2 XP), Survival 1 (2 XP).\index{Combat}\index{Stealth}\index{Survival}
\item Minor Talent: Iron Stomach (3 XP).\index{Iron Stomach}
\item \textbf{Total:} 32 XP (requires 1 bond or minor complication).\index{Bonds}\index{Complications}
\end{itemize}

\subsection{Mixed Build (32 XP)}
\label{subsec:mixed-example}
\index{Archetypes!Mixed!Example}

\begin{itemize}
\item Attributes: Body 2 (6 XP), Wits 2 (6 XP), Spirit 2 (6 XP).\index{Attributes}
\item Skills: Sway 2 (4 XP), Lore 1 (2 XP).\index{Sway}\index{Lore}
\item Minor Asset: Safehouse (4 XP).\index{Assets!Safehouse}
\item Cap 2 Follower (4 XP).\index{Followers!Cap 2}
\item \textbf{Total:} 32 XP.
\end{itemize}

\subsection{Mastermind Build (34 XP)}
\label{subsec:mastermind-example}
\index{Archetypes!Mastermind!Example}

\begin{itemize}
\item Attributes: Body 2 (6 XP), Wits 3 (9 XP), Spirit 2 (6 XP).
\item Skills: Sway 2 (4 XP).\index{Sway}
\item Cap 3 Follower (9 XP).\index{Followers!Cap 3}
\item \textbf{Total:} 34 XP (requires 2 bonds or complications).\index{Bonds}\index{Complications}
\end{itemize}

\section{Advancement Guidance}
\label{sec:advancement-guidance}
\index{Advancement!Guidance}

\begin{itemize}
\item Advancement should always reinforce narrative identity and character goals.\index{Narrative identity}
\item Consider how new abilities fit the character's story and relationships.\index{Relationships}
\item Balance immediate power with long-term narrative potential.\index{Power curve}
\item Work with the GM to ensure advancements make sense in the campaign context.\index{GM guidance}
\end{itemize}

\section{Optional Rule: Multi-Character Followers}
\label{sec:multi-character-followers}

\subsection*{Overview}
A player may control multiple characters by linking them as mutual Followers. This option also covers absent players' characters when consent is given.

\subsection*{Core Mechanics}
\begin{itemize}
  \item \textbf{Cap Tier +1:} A Follower may advance up to one Tier higher than their leader’s current Tier.
  \item \textbf{Multi-Character Play:} A player may designate two PCs as Followers of each other. Only one is the \emph{active} PC at a time; the other functions as a Follower.
  \item \textbf{Absent PCs:} With consent, an absent player’s PC may be run as a temporary Follower at Cap~Tier +1.
\end{itemize}

\subsection*{Switching Active Characters}
\label{sec:switching-active-characters}
\begin{itemize}
  \item \textbf{Switch Cost:} When switching, the previously active PC immediately becomes a Follower of the new active PC at Cap~Tier +1.  
  \item \textbf{Condition Gate:} A PC cannot be switched \emph{to} if Compromised, Seized, or otherwise narratively unfit for active duty.  
  \item \textbf{Narrative Justification:} The fiction must support the switch (returning from a journey, recovering from harm, taking point for a specific task).  
  \item \textbf{Strategic Choice:} Switching trades autonomy for support—gaining a powerful Follower but losing a full PC.  
\end{itemize}

\subsection*{Balance Notes}
\begin{itemize}
  \item Prevents abuse by making switching itself the cost.  
  \item Maintains narrative logic—leadership shifts naturally reshape party structure.  
  \item Creates strategic depth—switching is a meaningful tactical decision.  
  \item Keeps all characters relevant, encouraging upkeep even when not active.  
\end{itemize}