% =========================
% Fate's Edge — Magic System (Main)
% =========================
\chapter{Magic System}
\label{chap:magic-system}

\section{Design Philosophy}
\label{sec:magic-philosophy}

Magic in \textbf{Fate's Edge} is a dangerous negotiation with the fabric of reality. It is powerful and flexible—yet every attempt to shape it carries risk. Each die showing \textbf{1} generates \textbf{Story Beats (SB)}, which are prompts for twists and complications. The fiction leads; math follows.

\section{The Four Paths of Magic}
\label{sec:four-paths}

\subsection{Casting (Freeform Magic)}
\label{subsec:freeform-casting}

Freeform casting represents raw, improvisational magic.

\begin{description}
\item[Requirement] \textbf{Caster's Gift} Talent (2 XP).
\item[Mechanics] Two-step \emph{Weave \& Cast} colored by the Eight Elements; fiction-first targets and scope.
\item[Risk] Each 1 generates SB; DV scales with scope; backlash is expressed by Element (or its opposite).
\item[Limits] Heavy control effects (e.g., \texttt{[WARD]}, \texttt{[BANISH]}, \texttt{[UNWARD]}) require a printed source (Talent, \emph{Rite} text, or Spell result).
\end{description}

\subsection*{GM Guidance}\label{subsec:backlash-gm}
\begin{itemize}
\item \textbf{Show, then Nudge.} Lead with the fiction (smoke, fissures, whispers), then apply the smallest mechanical nudge that preserves drama.\index{Design Philosophy}
\item \textbf{Escalate with Consent.} Offer players the choice to escalate Minor to Major by taking +1 (SB) now to seize something they want.\index{Story Beats!player choice}
\item \textbf{One Bite per Cast.} Apply at most one backlash per cast/action unless a move explicitly stacks. Keep it punchy, not punitive.\index{Backlash!frequency}
\item \textbf{Clocks with Names.} Name clocks (\emph{Spreading Fire}, \emph{Inevitable Outcome}) so they feed the fiction and remind the table what's at stake.\index{Clocks}
\end{itemize}

\begin{tcolorbox}[title={Backlash Cheatsheet (margin-ready)},colback=gray!5,colframe=black]
\small \textbf{Minor = wobble, Major = lurch.} Types: Position/Eff, Clock +1/2 or +1, Condition, Cost. Offer Major by (SB)+1. Earth/Fate binds; Fire/Life burns/grows; Air/Luck scatters/ flips; Water/Obishaal pulls/opens.\par\smallskip
Place a mini version of Table~\ref{tab:backlash-8x2} on character sheets.
\end{tcolorbox}

\subsection{Rites Users (Runekeepers)}
\label{subsec:runekeepers}

Runekeepers bind themselves to structured \emph{Rites} from a single Patron.

\begin{description}
\item[Requirement] \textbf{Thiasos (Familiar)} (2 XP) \emph{and} \textbf{Codex} (4 XP). Runekeepers are restricted to \textbf{one} Patron.
\item[Mechanics] \textbf{Invoke} a known \emph{Rite} as \textbf{1 action}; on completion, mark \textbf{+1 Obligation} to that Patron. \emph{Push It} once/scene for amplified effect (\textbf{+1 Obligation}).
\item[Patron's Gift (Imbuement)] With Thiasos, once/scene as \textbf{1 action} imbue a held item for the scene with \textbf{+1 Melee} and \textbf{+1 Thematic} (a fixed Skill set by the Patron; see Table in \S\ref{sec:rites}). \emph{Push It} to extend one additional scene (\textbf{+1 Obligation}). A Codex is \emph{not} required for the Gift.
\end{description}

\section{Rites Difficulty Value (Expanded)}
\label{sec:rites-dv-expanded}
\index{Rites!Difficulty Value}
\index{DV}

\subsection*{Core Rule}

The Difficulty Value (DV) to cast a Rite is:

\[
\text{DV} = \max\!\big(\text{Obligation Cost} - \text{Spirit}, \, \text{Tier}\big)
\]

\begin{description}
  \item[Obligation Cost:] The Rite’s listed cost in Obligation segments. This reflects the Patron’s toll for the magic.
  \item[Spirit:] The caster’s Spirit attribute. Each point reduces the effective weight of the Obligation, representing inner resilience and willpower.
  \item[Tier:] The Rite’s intrinsic difficulty based on scope or potency. DV can never fall below this floor.
\end{description}

\subsection*{Design Philosophy}

This formula balances three forces:
\begin{itemize}
  \item \textbf{Debt vs. Strength:} Powerful Rites impose heavier Obligation, but high Spirit offsets that burden.
  \item \textbf{Narrative Tier Floor:} Even the simplest ritual of summoning or warding retains a minimum DV based on its Tier, ensuring gravity and consistency.
  \item \textbf{Scaling:} As characters grow, Spirit makes weaker Rites feel easier, while greater costs still challenge them.
\end{itemize}

\subsection*{Worked Examples}

\begin{itemize}
  \item \emph{Novice Example:} A Tier~1 Rite with Obligation Cost~2, cast by a Spirit~1 character:
  DV = $\max(2-1, 1) = 1$.
  The character feels only a slight strain, their Spirit covering most of the toll.

  \item \emph{Mid-Level Example:} A Tier~2 Rite with Obligation Cost~4, cast by a Spirit~2 character:
  DV = $\max(4-2, 2) = 2$.
  The Rite taxes them, but their Spirit prevents the burden from becoming overwhelming.

  \item \emph{Advanced Example:} A Tier~3 Rite with Obligation Cost~7, cast by a Spirit~3 character:
  DV = $\max(7-3, 3) = 4$.
  Here, Obligation dominates—the cost is heavy, and even strong Spirit cannot fully deflect it.
\end{itemize}

\subsection*{GM Guidance}

\begin{itemize}
  \item \textbf{Patron Themes:} Higher Obligation costs should reflect narrative weight, not just numbers. A Patron of decay may exact tolls in corruption, while a Patron of luck might demand reckless wagers.
  \item \textbf{Spirit as Fiction:} Encourage players to describe how their Spirit manifests—does it show as discipline, willpower, ritual focus, or raw charisma? Make the stat come alive in the fiction.
  \item \textbf{Scaling with Tier:} Remind players that no matter how trivial the fiction might feel, a Tier~2 or higher Rite is never “easy.” The floor maintains tension.
\end{itemize}

\section{Obligation Capacity}

A character’s \textbf{Obligation Capacity} equals Spirit + Presence.
Track total Obligation segments across all Patrons (or Symbols, for Invokers).

\begin{itemize}
  \item \textbf{Exceeding Capacity:} For each segment above Capacity, mark 1 Fatigue. The character cannot Invoke Rites or perform rituals until Obligation is reduced below Capacity.
  \item \textbf{Resolution:} Reduce Obligation through Downtime service, Patron tasks, ritual cleansing, or story resolution.
\end{itemize}

\textbf{Example:} Spirit~2 + Presence~3 = Capacity 5.
6 segments → Fatigue~1.
7 segments → Fatigue~2.
10 segments → Harm~1.
11 segments → Harm~2.

\subsection*{Optional Modules}

\begin{itemize}
  \item \textbf{Overflow DV:} If DV exceeds 5, consider applying minor narrative complications or Backlash risks on top of the roll, to show strain bleeding into the scene.
  \item \textbf{Patron Wrath Trigger:} A Patron may impose extra narrative tolls if the DV was reached primarily through Obligation rather than Tier—this shows over-dependence on their favor.
\end{itemize}

\subsection{Invokers (Symbol Path)}
\label{subsec:invokers}

Invokers use consecrated \textbf{Symbols} as ritual anchors to access a Patron's \emph{Rites} without a full bond.

\begin{description}
  \item[Requirement] \textbf{Patron's Symbol} (4 XP) per Patron; one Symbol per Patron. No Thiasos or Codex required.
  \item[Ritual Invocation] Perform the \emph{Rite} as a \textbf{ritual}: \textbf{DV + 1 rounds} to complete. Completion always marks \textbf{+1 Obligation}.
  \item[Symbol Benefit] When the correct \textbf{Symbol} is openly displayed/empowered for the Rite, reduce \textbf{DV by 1} (minimum 1) and reduce the Obligation cost by \textbf{1} (minimum 0).
  \item[No Symbol] You \emph{can} attempt a Patron's Rite without the Symbol, but you incur the following penalties: \textbf{+1 DV} \emph{(and therefore +1 round to cast)}, \textbf{+1 Obligation} \emph{(in addition to the base cost)}, and on \emph{Partial/Failure} the Keeper gains \textbf{+1 extra SB}. The Keeper may also worsen Position due to unstable focus.
  \item[Crack the Seal] Resolve a known Rite instantly as \textbf{1 action} by setting the Symbol to \textsc{Compromised} and marking \textbf{+2 Obligation} (\textbf{+3} if the Rite is High-Power). The GM may immediately spend 1 on-theme SB. Restore the Symbol in Downtime (test DV 3 or via fiction) or spend 1 XP.
  \item[Limits] Symbols must be openly displayed during the ritual. Carrying \textbf{4+} Symbols causes \textbf{+1 Obligation} on the first ritual each scene. Rival or hostile Symbols in the scene may impose \textbf{--1 Position} and \textbf{+1 Obligation} at the Keeper's discretion.
\end{description}

\subsection*{Borrowed Grace}
\label{talent:borrowed-grace}
\index{Talents!Invoker}\index{Imbuement!Lesser}

\textbf{Type:} Invoker Talent — \textit{Lesser Imbuement}

\subsubsection*{Use}
\begin{itemize}
  \item \textbf{Cost:} \textbf{1 Boon}, \textbf{1 action}.
  \item \textbf{Effect (pick one on use):} \textbf{+1 Melee} \emph{or} \textbf{+1 Thematic} (your table's signature/thematic Skill).
  \item \textbf{Duration:} \textit{Single action/attack} (instantaneous boost only).
  \item \textbf{Requirement:} Wield/display the relevant Patron's \textbf{Symbol}.
  \item \textbf{Obligation:} Immediately mark \textbf{+1 Obligation} to that Patron (see \S\ref{sec:obligation}).
  \item \textbf{Limits:} Cannot be extended, stacked, or \emph{Pushed} for duration. Using \emph{Borrowed Grace} while the Symbol is \textsc{Compromised} lowers your \textbf{Position} by one step \emph{(or imposes \(-1\) die if already \textbf{Desperate})}.)
\end{itemize}

\subsection*{Fictional Framing}
A quick, rule-bending channel through a Patron's \emph{Symbol}—a sliver of grace, borrowed for a moment and paid for in debt.

\subsection*{Table Guidance (1-liners)}
\begin{itemize}
  \item \textbf{Combat:} Spike a strike vs. a tough foe; or steady a parry in a desperate bind.
  \item \textbf{Skill:} Nudge a pivotal social/ritual/track roll tied to the Patron's sphere.
  \item \textbf{Fallout:} Repeated use accrues \textbf{Obligation}; NPC faithful may notice "stolen" grace.
\end{itemize}

\subsection*{Balance Notes}
\begin{itemize}
  \item Weaker than full Imbuement: \emph{one} action, no sustain, upfront Obligation.
  \item \textbf{Symbol dependency:} No Symbol, no channel (concealed or lost Symbol = no effect).
\end{itemize}

\subsection*{GM Hooks (quick picks)}
\begin{itemize}
  \item \textbf{Compel Debt:} A Patron agent arrives when Obligation crosses a tick.
  \item \textbf{Clash of Signs:} Using rival Symbols back-to-back risks minor \textbf{Backlash} (drop Position or +1 SB).
  \item \textbf{Spotlight Tell:} Brief visual tell (scent, sigil flare) marks the borrowing to observant NPCs.
\end{itemize}

%--- Canonical Leash Capacity rule (drop-in replacement) ----------------------

\begin{fatebox}[Summoning Core Mechanics]
  \begin{tabularx}{\textwidth}{lX}
  \toprule
  \textbf{Mechanic} & \textbf{Description and Requirements} \\
  \midrule
  \textbf{Call} & 1 Action to manifest the spirit at \textit{Near} range; choose a Spirit Template aligned to fiction or Patron domain. \\
  \textbf{Bind} & Spend 1 Boon \emph{or} mark 1 Fatigue to establish initial control. \\
  \textbf{Leash Capacity} & Set equal to \textit{Cap} $+$ \textit{Spirit} \emph{segments}.  
  (\textit{Cap} is the Outsider's tier: Cap~1 for Lesser, Cap~3 for Greater.) \\
  \textbf{Tick Leash} & Whenever the spirit takes Harm, you command it against its nature, you split focus, a rival contests it, it moves \textit{Close $\rightarrow$ Far} rapidly, or crosses a \texttt{[WARD]} (\textit{DV = Cap}). \\
  \textbf{Departure} & When the Leash fills, the spirit acts to its nature once, then departs (or turns hostile at GM discretion). \\
  \bottomrule
  \end{tabularx}
  \end{fatebox}
  
  \paragraph{Procedure.}
  \begin{enumerate}
    \item \textbf{Call (1 Action):} A spirit manifests at \textit{Near}. Choose a Spirit Template appropriate to the scene or Patron.
    \item \textbf{Bind:} Spend 1 Boon \emph{or} mark 1 Fatigue to anchor the connection.
    \item \textbf{Leash Capacity:} Record Leash Capacity $=$ \textit{Cap} $+$ \textit{Spirit} \emph{segments}. Draw a clock to track strain.
    \item \textbf{Command:} Each round, issuing a meaningful order uses your Action. Commands contrary to the spirit's nature tick the Leash.
    \item \textbf{Maintain:} If you split focus or perform other significant actions while it acts on your order, tick the Leash.
    \item \textbf{Departure:} When the Leash fills, the spirit acts to its nature once, then departs. Use this to escalate or reveal consequences.
  \end{enumerate}
  
  \paragraph{Example.}
  \textit{Kestra calls a Cap~3 fire elemental to aid in battle. She spends 1 Boon to Bind it.  
  The elemental's Leash Capacity is 7 segments (\textit{Cap}~3 $+$ \textit{Spirit}~4). When it takes Harm, the GM ticks the Leash. Later, Kestra splits focus to issue orders while attacking, ticking again.  
  Careful management and Boon Finesse keep the bond stable—until the elemental's fury tests her will. After the battle, she marks her Spirit Bond Clock +1 for the shared victory.}
\section{The Nature of Magic}
\label{sec:nature-of-magic}

\begin{itemize}
\item \textbf{Volatile by design:} Each working pushes boundaries that resist being bent.
\item \textbf{Risk embodied:} Each 1 on any magic roll generates SB the GM can spend for backlash or twists.
\item \textbf{Narrative weight:} Every magical action alters the scene, even on a success.
\item \textbf{Thematic consequence:} Backlash aligns with the invoked Element or its opposition.
\end{itemize}

\section{The Eight Elements}
\label{sec:eight-elements}

\subsection{Physical}
\paragraph{Earth} Solidity, structure; shape/sense/move stone; backlash: rigidity/collapse. \quad
\paragraph{Fire} Energy, change; ignite/heat/purify; backlash: spread/scorch. \quad
\paragraph{Air} Motion, sound; push/pull/resonance; backlash: dispersal/whip. \quad
\paragraph{Water} Flow, repair; channel/cleanse/mend; backlash: flood/contaminate.

\subsection{Metaphysical}
\paragraph{Fate} Causality, oaths, anti-magic; backlash: paradox/closure. \quad
\paragraph{Life} Vitality, growth, repair; backlash: overgrowth/fever. \quad
\paragraph{Luck} Chance, openings; backlash: side-coincidence/irony. \quad
\paragraph{Death/Dreams (Obishaal)} Thresholds, Ways Between; backlash: thin walls/nightmares.

\section{Magical Arts}
\label{sec:magical-arts}

Define your \emph{Art} (gesture/medium, two typical Elements, signature style). If the Art is clearly honored in fiction, gain \textbf{+1 die} on your \textbf{Cast} once/scene (counts toward +3 cap). Working hard against your Art can worsen Position or pre-load backlash on a Partial.

\section{Casting Loop (Freeform)}
\label{sec:casting-loop}

\textbf{Channel:} Focus and draw Potential (e.g., Wits+Arcana); successes become shaping fuel; each 1 generates SB.\\
\textbf{Weave:} Next turn, shape the effect (e.g., Wits+Art); apply the Description Ladder (Basic/Detailed/Intricate) per core rules.\\
\textbf{Backlash:} GM spends SB thematically by Element; severity scales with SB and scope. Boons do not reduce SB unless a source says so.

\section{Magic in Combat}
\label{sec:magic-combat}

Casting typically takes two actions (Channel + Weave). Runekeeper \emph{Rites} resolve in one action (with Obligation risk). Invoker rituals are usually too slow for a fight—use \emph{Crack the Seal} for instant results at high cost. \texttt{[COUNTER]} can interrupt any magical action in its window.

\section{Path Comparison}
\label{sec:path-comparison}

\begin{table}[htbp]
\centering
\begin{tabular}{p{3.1cm}p{4cm}p{4cm}p{4cm}p{4cm}}
\toprule
\textbf{Aspect} & \textbf{Caster (Freeform)} & \textbf{Runekeeper (Rites)} & \textbf{Invoker (Symbols)} & \textbf{Summoner (Pact)} \\
\midrule
Access Cost & Caster's Gift (2 XP) & Thiasos + Codex (6 XP) & Symbol (4 XP per Patron) & Lesser/Greater Pactwright \\
Speed & Medium (2 actions) & Fast (1 action) & Slow (ritual) / Fast w/ Seal & Fast (Call = 1 action) \\
Risk Type & SB backlash (Elemental) & Obligation (Patron ledger) & Symbol compromise + Obligation & Leash fill + command costs \\
Breadth & High (fiction-gated) & Medium (defined \emph{Rites}) & Medium (breadth across Patrons) & Medium (by Templates/Cap) \\
Sustain & Fatigue/backsplash & Obligation; Push adds +1 & Obligation; Symbol state gates & Leash ticks; Boon Finesse \\
\bottomrule
\end{tabular}
\caption{Comparison of Magic Paths}
\end{table}

\section{Guardrails}
\label{sec:magic-guardrails}

\begin{itemize}
\item \textbf{Duration defaults:} Buffs $\approx$ 3 beats; areas 1 beat. Sustaining costs 1 Fatigue/beat.
\item \textbf{Stacking:} Same-source effects do not stack; take the best instance.
\item \textbf{Assist cap:} +3 dice total from assists/buffs.
\item \textbf{Over-Stack:} Active magic can count as structural advantages for Over-Stack.
\item \textbf{Plausibility:} All effects must fit the fiction and established limits.
\end{itemize}

% Patrons & Rites
\input{patrons/index.tex}

% --- Fate's Edge SRD — Section 6: Summons & Outsiders ---
% Include this file from your main .tex with: \input{06-summons.tex}
\floatbarrier
\clearpage

\section{Summons and Outsiders}

\subsection{Definition}
An \textbf{Outsider} is any being not native to the world of Fate's Edge. This includes summoned spirits, demons, celestials, and entities that arrive from beyond the veil of the Eight Elements. They are powerful but dangerous to bind.

\subsection{Summoning (Pact-Whisperer Core)}
Summoning is a way to call and bind Outsiders for temporary aid.

\begin{enumerate}
  \item \textbf{Call} (1 Action): A spirit manifests at Near range. Choose a Spirit Template.
  \item \textbf{Bind}: Choose one: spend 1 Boon or mark 1 Fatigue.
  \item \textbf{Leash}: Set Leash = Cap + Command segments (Cap is the Outsider's tier, typically 1/3/5 for Lesser/Greater/Elder).
  \item \textbf{Tick Leash} whenever any occur:
    \begin{itemize}
      \item Spirit takes harm.
      \item You command against its nature.
      \item You split focus (take another significant action while it acts).
      \item A rival contests it.
      \item It moves from Close to Far quickly.
      \item It crosses a [WARD].
    \end{itemize}
  \item \textbf{Departure}: When the Leash fills, the spirit acts to its nature once, then departs.
\end{enumerate}

\textbf{Limits:} Only one active summoned spirit at a time (unless a Talent says otherwise). All summons depart at Downtime unless explicitly sustained.

\subsection{Boon Finesse}
Once per round, you may spend 1 Boon to clear 1 tick from your current spirit's Leash. You cannot do this after the Leash has filled.

\subsection{Outsider Caps}
\begin{itemize}
  \item PC-summoned Outsiders: Cap is limited by Talents (Lesser = 1, Greater = 3).
  \item NPC Outsiders: GM assigns based on story needs (Lesser = 1, Greater = 3, Elder = 5).
\end{itemize}

\subsection{Tags for Summons \& Outsiders}
Certain Tags specifically interact with Outsiders.

\begin{description}[leftmargin=1.5em, style=nextline]
  \item[WARD:] Creates a magical edge/zone that Outsiders must test to cross.
    \begin{itemize}
      \item DV = Outsider's Cap.
      \item Hit: Outsider crosses and its Leash gains +DV segments.
      \item Partial: Outsider crosses and its Leash gains +1 segment.
      \item Miss: Outsider fails to cross this beat.
    \end{itemize}
  \item[BANISH:] Drives a visible Outsider toward departure.
    \begin{itemize}
      \item DV = Outsider's Cap.
      \item Hit: Add +DV segments to its Leash (or Exit Tally).
      \item Partial: Add +1 segment.
      \item Miss: No effect.
    \end{itemize}
  \item[UNWARD:] Suppresses or dismisses a [WARD].
    \begin{itemize}
      \item DV by fiction (materials, sanctity, prep, locus, opposition).
      \item Hit: Ward dismissed/suppressed.
      \item Partial: Ward suppressed briefly (1 beat).
      \item Miss: No effect.
    \end{itemize}
\end{description}

\subsection{Unified Leash / Exit Tally System}
\begin{itemize}
  \item Summoned Outsiders track their service via a \textbf{Leash} (Cap + 2 segments).
  \item Non-summoned Outsiders affected by [WARD] or [BANISH] gain a temporary \textbf{Exit Tally} = Cap + 2. When the tally fills, they act to nature once, then depart.
\end{itemize}

\subsection{GM Guidance}
\begin{itemize}
  \item Summons are not permanent allies; they are volatile forces.
  \item Always color Outsider behavior by their Elemental resonance and domain.
  \item When the Leash fills, deliver a memorable "act to nature" moment before they vanish.
  \item Use SB to escalate Outsider complications: a jealous Patron, a backlash of strange omens, or collateral spiritual harm.
\end{itemize}

% !TEX root = srd_main.tex
% SRD Insert: Elemental Backlash Tables — 8 Elements × Minor/Major
% Assumes booktabs, tabularx, xcolor, tcolorbox are available

\section{Elemental Backlash}\label{sec:backlash-tables}
\index{Backlash}\index{Elements}\index{Story Beats}\index{Realms}\index{Obishaal@Obishaal (Dreams/Thresholds)}

When magic disturbs the weave, the world pushes back. Backlash manifests as fiction-first complications with light mechanical teeth. Each element (and its metaphysical counterpart) has a \textbf{Minor} and \textbf{Major} pattern.

\begin{tcolorbox}[title={Using Backlash at the Table},colback=gray!5,colframe=black]
\textbf{Trigger.} A roll shows a 1 (gaining a (SB)) or the text explicitly says "accept 1 (SB) to escalate."\newline
\textbf{Choose One:} Apply the table's Minor effect, or escalate to Major by adding $\mathbf{+1}$ (SB) immediately.\newline
\textbf{Mechanical Nudge Types.} \emph{Position/Effect shift}, \emph{Clock tick (1/2)}, \emph{Condition}, or \emph{Immediate Cost}.\index{Story Beats!backlash escalation}
\end{tcolorbox}

\subsection*{Realms and Counterparts}\label{subsec:realms-counterparts}
\begin{itemize}
\item \textbf{Earth} \emph{(Realm: Stone)} $\leftrightarrow$ \textbf{Fate} \emph{(Anti-magic, inevitability)}\index{Earth}\index{Fate}
\item \textbf{Fire} \emph{(Realm: Ember)} $\leftrightarrow$ \textbf{Life} \emph{(Vital spark, growth)}\index{Fire}\index{Life}
\item \textbf{Air} \emph{(Realm: Gale)} $\leftrightarrow$ \textbf{Luck/Fortune} \emph{(Ephemera, unlikely turns)}\index{Air}\index{Luck}\index{Fortune}
\item \textbf{Water} \emph{(Realm: Tides)} $\leftrightarrow$ \textbf{Death/Dreams/Thresholds (Obishaal)} \emph{(Passage, veils, the Ways Between)}\index{Water}\index{Death}\index{Dreams}\index{Thresholds}\index{Obishaal@Obishaal}
\end{itemize}

\vspace{0.5em}

% ========================= MASTER TABLE =========================
\begin{table}[h]
\centering
\caption{Backlash by Element (Minor / Major)}
\label{tab:backlash-8x2}
\renewcommand{\arraystretch}{1.12}
\begin{tabularx}{\linewidth}{>{\bfseries}l l >{\raggedright}X >{\raggedright}X}
\toprule
Element & Realm / Counterpart & Minor Backlash (fiction • nudge) & Major Backlash (fiction • nudge) \\
\midrule
Earth & Stone / \emph{Fate} & \emph{Stone binds.} Dust cakes tools; footing slips. • \textbf{Position},–1 or mark \textsc{Encumbered} (light). & \emph{Ground claims.} A fissure opens or masonry seizes. • \textbf{Clock},+1/2 on \emph{Collapse/Entrap} or \textbf{Condition},: \textsc{Pinned}. \\
Fire & Ember / \emph{Life} & \emph{Heat flares.} Smoke blinds; sparks bite. • \textbf{Effect},–1 or \textbf{Condition},: \textsc{Singed} (disadvantage to precise actions). & \emph{Blaze takes.} Fuel ignites or blood races fever-hot. • \textbf{Clock},+1 on \emph{Spreading Fire} or take \emph{1 Harm} ignoring armor. \\
Air & Gale / \emph{Luck} & \emph{Winds misplace.} Words scatter; aim wavers. • \textbf{Position},–1 or \textbf{Clock},+1/2 on \emph{Alarmed Attention}. & \emph{Fickle turn.} An unlikely mishap hits you instead. • \textbf{Immediate Cost},: lose a tool/use \emph{or} \textbf{(SB)},+1 as chaos compounds. \\
Water & Tides / \emph{Obishaal} & \emph{Seep and swell.} Grip slicks; tide shifts. • \textbf{Effect},–1 or \textbf{Condition},: \textsc{Waterlogged} (gear slow). & \emph{Undertow calls.} Passage veers; something is pulled through. • \textbf{Clock},+1 on \emph{Flood/Wayward Current} \emph{or} introduce a \emph{brief} intrusion from the Ways Between. \\
Fate & Anti-magic / \emph{Earth} & \emph{Lines harden.} Probability resists change. • \textbf{Effect},–1 on overt magic or \textbf{Clock},+1/2 on \emph{Inevitable Outcome}. & \emph{Edict falls.} A declared cost must be paid now. • \textbf{Immediate Cost},: sacrifice a resource \emph{or} \textbf{(SB)},+1 and mark \textsc{Omen}. \\
Life & Vital spark / \emph{Fire} & \emph{Growth misfires.} Vines choke; pulse surges. • \textbf{Condition},: \textsc{Overgrowth} (tethered) or \textbf{Effect},–1 on precision. & \emph{Riot of life.} Parasites bloom; healing twists. • \textbf{Clock},+1 on \emph{Biohazard} \emph{or} convert 1 healing into \textbf{(SB)},+1. \\
Luck & Ephemera / \emph{Air} & \emph{Odds flip.} A near-sure thing slips. • \textbf{Position},–1 or \textbf{Clock},+1/2 on \emph{Unwelcome Coincidence}. & \emph{Snake eyes.} Catastrophic fluke. • Force a \emph{re-roll}; if any 1 appears, \textbf{(SB)},+1 \& apply Minor again. \\
Death/Dreams & Ways Between / \emph{Water} & \emph{Veil thins.} Echoes, whispers, cold breath. • \textbf{Condition},: \textsc{Shaken} (first action –1) or \textbf{Clock},+1/2 on \emph{Haunting}. & \emph{Threshold opens.} A path misaligns; a revenant claims a due. • \textbf{Clock},+1 on \emph{Crossing Due} \emph{or} immediate scene intrusion from Obishaal. \\
\bottomrule
\end{tabularx}
\end{table}

% !TEX root = resource_guide_main.tex
% Extended Chapter: Rituals — Philosophy, Procedures, Costs, and Advanced Use
% Assumes: booktabs, tabularx, xcolor, tcolorbox, enumitem, hyperref, amsmath

\section{Rituals (Extended)}\label{sec:rituals-extended}
\index{Rituals}\index{Story Beats}\index{Backlash}\index{Elements}\index{Realms}\index{Obishaal@Obishaal}\index{Patrons}\index{Symbols}\index{Runekeeper}\index{Invoker}\index{Summoner}\index{Caster}

Rituals are \emph{slow magic}: explicit intent, staged action, and negotiated risk. They are how characters bend the world carefully, trading time, components, and narrative exposure for precise results. This section expands the SRD quick-start (\S\ref{sec:universal-rituals}) with procedures, dials, and worked examples.

\subsection{Design Goals}\label{subsec:ritual-goals}\index{Design Philosophy}
\begin{itemize}
\item \textbf{Fiction-first.} Components and steps are story handles, not inventory chores.
\item \textbf{Visible costs.} Every ritual declares \emph{what it costs} (time, component loss, conditions) and \emph{how it risks} (SB)/backlash.
\item \textbf{Tempting choices.} Players can \emph{push}—accept (SB) to escalate position/effect/scale.
\item \textbf{Portable.} Works for Runekeepers, Invokers, Summoners, and Free Casters with minimal chassis-specific tweaks.
\end{itemize}

\subsection{Ritual Procedure}\label{subsec:ritual-procedure}
Use this skeleton for any ritual, published or improvised.

\begin{enumerate}[label=\textbf{Step \arabic*:}, leftmargin=2.2em]
\item \textbf{State Intent.} What do you want? Clarify element/Realm if obvious (Fire for heat, Water for memory, Fate for anti-magic, Obishaal for thresholds).\index{Elements!choosing}
\item \textbf{Choose Scope.} Size, duration, range, and detail. Start modest; escalations come later.
\item \textbf{Lay Components.} Name \emph{two things}: (a) \textbf{Focus} (tool/site/patron sign), (b) \textbf{Fuel} (herb, blood, pact, memory). Decide which is consumed vs. retained.\index{Components}
\item \textbf{Set Time.} Default: \emph{Low 1 minute / Med 5–10 minutes / High 15–20 minutes}. More time improves position/effect; rushing worsens it.
\item \textbf{Call Risks.} Point to the element's \textbf{Backlash} (\S\ref{sec:backlash-condensed}) and the default \textbf{(SB) trigger}: any 1 rolled creates a (SB); re-rolling 1s does not remove (SB) and may add another.
\item \textbf{Roll and Resolve.} Apply position/effect and any clocks. Offer a \emph{push}: take +1 (SB) to step up result now.
\item \textbf{Mark Costs.} Consume components, apply Conditions, or tick wear/concurrency (per chassis). Close the scene hooks the ritual created.
\end{enumerate}

\subsection{Component Economy}\label{subsec:component-economy}
Components are levers, not taxes. Use them to signal tone and stakes.

\begin{table}[h]
\centering
\caption{Components as Narrative Levers}
\label{tab:ritual-components}
\renewcommand{\arraystretch}{1.12}
\begin{tabularx}{\linewidth}{>{\bfseries}l X X}
\toprule
Type & Examples & Mechanical Nudge \\
\midrule
Focus (retained) & Patron token, true-name sigil, saint's nail, ancestral blade. & +1 \emph{position} on setup or advantage to related follow-up actions. \\
Fuel (consumed) & Herb bundle, salt vial, blood drop, silver bead, memory-laden note. & –1 \emph{time step} (faster) or +1 \emph{effect} step if expensive/rare. \\
Site (context) & Crossroads, standing stones, bathhouse, bell tower at midnight. & Shift backlash element or re-route it (e.g., into \emph{Alarmed Attention} clock). \\
Vow (social) & Sworn phrase, bargain pledge, offered favor. & If broken, immediate (SB) +1 and an intrusion linked to the vow. \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Teamwork and Aid}\label{subsec:ritual-teamwork}\index{Teamwork}
\begin{itemize}
\item \textbf{Hands \& Voices.} Each assistant names one component they contribute; either reduces cast time \emph{or} accepts up to 1 (SB) on the caster's behalf once per ritual.
\item \textbf{Focus Chain.} Passing the Focus around the circle grants advantage on the finishing action but risks \textsc{Distracted} if interrupted.
\item \textbf{Distributed Load.} Splitting a High ritual into two coordinated Mediums avoids a Major backlash but creates two Minor hooks instead.
\end{itemize}

\subsection{Clocks and Outcomes}\label{subsec:ritual-clocks}\index{Clocks}
Tie every consequential ritual to \textbf{named clocks}. Examples: \emph{Spreading Fire}, \emph{Inevitable Outcome}, \emph{Crossing Due}, \emph{Alarmed Attention}. Advancing or reducing clocks is often better than flat bonuses.

\begin{tcolorbox}[title={Outcome Palette},colback=gray!5,colframe=black]
\small \textbf{On a strong result:} full effect, +1 effect step, or Clock –1.\newline
\textbf{On a mixed:} effect with a cost (component consumed; condition applied).\newline
\textbf{On a weak:} effect limited; Clock +1/2; Minor Backlash.\newline
\textbf{On a push:} player may take (SB) +1 to upgrade one step immediately.\end{tcolorbox}

\subsection{Backlash Integration}\label{subsec:ritual-backlash}
Use the condensed table (\S\ref{tab:backlash-condensed}). Calibrate by scene weight: exploratory scenes favor Minor; pivotal moments bait a Major via (SB) +1.

\subsection{Chassis-Specific Notes}\label{subsec:ritual-chassis}
\paragraph{Runekeeper.}\index{Runekeeper} Embed Rites as \emph{accelerants}: a published Rite may count as a Focus that upgrades position or halves time.
\paragraph{Invoker.}\index{Invoker} Symbols accumulate \emph{wear}; a maintenance rite can clear 1 wear mid-scene on success \emph{or} shift backlash from Major to Minor.
\paragraph{Summoner.}\index{Summoner} \emph{Gate} effects occupy concurrency slots. Disruption on broken terms: (SB) +1 and the entity acts on its last instruction.
\paragraph{Caster (Free).}\index{Caster} Tags become explicit ritual steps (\emph{bind, veil, reveal})—chain two compatible tags once/scene for a synergy bump without extra cost.

\subsection{Ritual Templates}\label{subsec:ritual-templates}
Use these fill-in cards to author new content quickly.

\begin{tcolorbox}[title={Template: Utility Rite (Low)},colback=gray!3,colframe=black]
\textbf{Name}: \rule{0.6\linewidth}{0.4pt} \quad \textbf{Element}: \rule{0.28\linewidth}{0.4pt}\\
\textbf{Cast Time}: 1 minute \quad \textbf{Scope}: pocket-scale\\
\textbf{Components}: Focus (kept): \rule{0.4\linewidth}{0.4pt}; Fuel (consumed): \rule{0.35\linewidth}{0.4pt}\\
\textbf{Effect}: \rule{0.9\linewidth}{0.4pt}\\
\textbf{Cost}: \rule{0.9\linewidth}{0.4pt}\\
\textbf{Backlash}: Minor (\S\ref{sec:backlash-condensed}). \textbf{Push}: take (SB) +1 to upgrade one step.
\end{tcolorbox}

\begin{tcolorbox}[title={Template: Scene Rite (Med)},colback=gray!3,colframe=black]
\textbf{Name}: \rule{0.6\linewidth}{0.4pt} \quad \textbf{Element}: \rule{0.28\linewidth}{0.4pt}\\
\textbf{Cast Time}: 5–10 minutes \quad \textbf{Scope}: room/street\\
\textbf{Components}: Focus (kept): \rule{0.4\linewidth}{0.4pt}; Fuel (consumed): \rule{0.35\linewidth}{0.4pt}; Site: \rule{0.3\linewidth}{0.4pt}\\
\textbf{Effect}: \rule{0.9\linewidth}{0.4pt}\\
\textbf{Cost}: \rule{0.9\linewidth}{0.4pt}\\
\textbf{Backlash}: Minor; offer Major via (SB) +1. \textbf{Clocks}: \rule{0.5\linewidth}{0.4pt}
\end{tcolorbox}

\begin{tcolorbox}[title={Template: Set-Piece Rite (High)},colback=gray!3,colframe=black]
\textbf{Name}: \rule{0.6\linewidth}{0.4pt} \quad \textbf{Element}: \rule{0.28\linewidth}{0.4pt}\\
\textbf{Cast Time}: 15–20 minutes \quad \textbf{Scope}: block/fort\\
\textbf{Components}: Focus (kept): \rule{0.4\linewidth}{0.4pt}; Fuel (consumed): \rule{0.35\linewidth}{0.4pt}; Site: \rule{0.3\linewidth}{0.4pt}; Vow: \rule{0.3\linewidth}{0.4pt}\\
\textbf{Effect}: \rule{0.9\linewidth}{0.4pt}\\
\textbf{Cost}: \rule{0.9\linewidth}{0.4pt}\\
\textbf{Backlash}: Likely Major; bait with (SB) +1. \textbf{Clocks}: \rule{0.5\linewidth}{0.4pt}
\end{tcolorbox}

\subsection{Worked Examples}\label{subsec:ritual-examples}
\paragraph{Example 1: Quiet Veil (Team Infiltration).}\index{Rituals!examples}
\emph{Intent:} silence the group for one scene. \emph{Scope:} corridor sweep. \emph{Components:} ash (fuel, consumed), bell (focus, kept). \emph{Time:} 5 minutes. \emph{Risks:} Air/Luck Minor on 1; offer Major to avoid dogs' scent. \emph{Roll:} mixed—effect with cost. \emph{Outcome:} \textsc{Muted} condition until scene ends; Clock –1/2 on \emph{Patrol Pass}. Player takes (SB) +1 to also foil scent (Major avoided by paying the (SB)).

\paragraph{Example 2: River's Memory (Investigation).}
\emph{Intent:} view last night's ferry landing. \emph{Scope:} a few minutes of blurred images. \emph{Components:} bowl, token from dock. \emph{Time:} 10 minutes. \emph{Risk:} Water/Obishaal Minor. \emph{Roll:} strong—clear image; token ruined per Cost. \emph{Outcome:} Clock –1 on \emph{Where did the courier go?}; whisper from the Ways foreshadows a revenant (hook).

\paragraph{Example 3: Fate-Splice (Boss Rescue).}
\emph{Intent:} move the poison consequence from the prince to the knight. \emph{Scope:} one Major consequence. \emph{Components:} paired names on vellum; vow. \emph{Time:} 15 minutes. \emph{Risk:} Fate/Earth. \emph{Roll:} weak—Minor Backlash; \emph{Inevitable Outcome} +1/2. \emph{Push:} (SB) +1 to capture the full consequence anyway. \emph{Outcome:} knight bears the poison; the \textsc{Omen} mark appears (future hook).

\subsection{Safety and Consent}\label{subsec:ritual-safety}\index{Safety}
Rituals often touch body horror, spiritual intrusion, or coercive bargains. Use lines/veils, X-card, script change, or your table's preferred tools. Make \textbf{vows} opt-in; provide non-coercive alternatives with different trade-offs.

\subsection{Optional Modules}\label{subsec:ritual-modules}
\paragraph{Entropy Counters.} Track ritual entropy per scene; at 3+ entropy, the next Minor backlash upgrades to Major automatically. Resets on scene change.
\paragraph{Resonance Sites.} Mark places that boost one element (+1 effect) and hinder its counterpart (–1 position). Crossing a resonance flips the pairing.
\paragraph{Material Tags.} Let special materials act like tags (cold iron, voidglass) granting narrow immunities or redirecting backlash type.

\subsection{GM Troubleshooting}\label{subsec:ritual-troubleshooting}
\begin{itemize}
\item \textbf{Pacing drifts long.} Shorten cast time by consuming an extra Fuel component; keep one meaningful step.
\item \textbf{Risk feels toothless.} Name a clock and advance it on mixed/weak even if the effect lands.
\item \textbf{Runekeeper dominates.} Insert a \emph{Rune Draw} tell for high-grade rites \emph{or} grant Invokers a mid-scene maintenance clear on a solid success.
\item \textbf{Summons flood scene.} Enforce concurrency and Disruption (\S\ref{subsec:ritual-chassis}).
\item \textbf{Casters feel mushy.} Require two explicit tags per ritual step (bind/veil/reveal); grant once/scene synergy bump.
\end{itemize}

\begin{tcolorbox}[title={Summary},colback=gray!5,colframe=black]
Rituals trade \textbf{time, components, and exposure} for \textbf{precision and scale}. Keep costs visible, risk tempting, and outcomes named via clocks. Offer players the choice to buy bigger results with (SB)—then pay off every hook
\end{tcolorbox}


%----------------------------------------
\section{Summoning (Pact-Whisperer) — GM Mechanics}
\label{subsec:summoning-gm}

\begin{fatebox}[Core]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Step} & \textbf{Rule (GM-Facing)} \\
\midrule
\textbf{Access} & Requires \textbf{Pact-Whisperer} (2 XP). Talents gate capacity: \textbf{Lesser Pactwright} (Cap~1), \textbf{Greater Pactwright} (Cap~3). With both, one of each may be maintained. \\
\textbf{Call} & \emph{1 Action}. Manifest at \textit{Near}. Choose a fitting Spirit Template (scene/Patron aligned). \\
\textbf{Bind} & Pay 1 Boon \emph{or} mark 1 Fatigue. \\
\textbf{Leash Capacity} & Set \(\textbf{Leash} = \textit{Cap} + \textit{Spirit}\) \emph{segments}. (Cap~1 = Lesser, Cap~3 = Greater.) \\
\textbf{Tick Triggers} & Tick on: spirit takes Harm; command vs.\ nature; summoner performs separate concentration-requiring action while commanding; rival contests; crossing a \texttt{[WARD]} successfully (test \(\mathrm{DV}=\mathrm{Cap}\)). \\
\textbf{Act/Order} & A meaningful new order uses the summoner's Action; \emph{Quick Commands} (attack nearest, hold doorway, relocate within \textit{Near}, fetch and return) do not. \\
\textbf{Departure} & When Leash fills, spirit acts to its nature once, then departs (or turns hostile at GM discretion). \\
\bottomrule
\end{tabularx}
\end{fatebox}

\paragraph*{Spirit Bond (Light Progress).}
Track a \textbf{Spirit Bond Clock [4]} per recurring spirit type. Mark on shared victories, good handling, or mutual aid. At 2: +1 die to communicate. At 4: on natural departure, +1 Boon and spirit becomes \textbf{Favored} (its Leash $-1$). \textit{Near-Miss:} once/session per type, mark +1 on a meaningful failed Call/Bind.

\paragraph*{Specializations (Pick When Relevant).}
Combat Specialist (+1 Harm melee; ignore first Harm on attacks) \quad
Scout Form (stealth/range; carry up to Cap~1: 2\,kg, Cap~3: 10\,kg; dragging $\le 3\times$; overburden = tick) \quad
Utility Spirit (simple tasks) \quad
Shield Guardian (interpose; convert Harm\(\rightarrow\)Fatigue).

\paragraph*{Economy \& Limits.}
\begin{itemize}
  \item \textbf{Boon Finesse:} Once/round, spend 1 Boon to clear 1 tick (before fill).
  \item \textbf{Order of Action:} Spirit acts immediately after the command is given.
  \item \textbf{Concurrency:} One active spirit unless a Talent states otherwise; excess costs 1 Fatigue per extra Cap point.
  \item \textbf{End-State:} All summons end at Downtime unless sustained by Rite/Asset.
\end{itemize}

\paragraph*{GM Reminders (No New Clocks).}
Use only the \textbf{Leash} and optional \textbf{Spirit Bond} clocks. \emph{Quick Commands} should be crisp rulings; tick when the player overreaches the spirit's nature or fiction. Clarify that "splitting focus" means performing a separate, concentration-requiring action while actively directing the spirit.

%----------------------------------------
\section{Cantor's Path (Songs) — GM Mechanics}
\label{talent:cantors-path-gm}

\begin{fatebox}[Core]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Element} & \textbf{Rule (GM-Facing)} \\
\midrule
\textbf{Access} & Talent: \textbf{Cantor's Path} (8 XP). Prereqs: Lore~1+, Performance~2+, Presence~2+. \\
\textbf{Scope} & \textbf{Low Rites as Songs} only (counts as knowing for performance use). \\
\textbf{Cast Test} & \emph{Lore + Performance vs.\ DV} (typical DV~2–3). \\
\textbf{Timing} & Start with 1 Action; resolves at \emph{start of next turn} unless \emph{Pushed}. \\
\textbf{Cost} & Pay listed materials. \emph{No Obligation} on success. \\
\textbf{Visibility} & Songs are noticeable; on Failure or Push, assume observers take note. \\
\bottomrule
\end{tabularx}
\end{fatebox}

\paragraph*{Corruption (Light).}
Track a \textbf{Corruption Clock} with \(\text{segments}=\) \textbf{Body}. Mark when: \emph{Push}, perform a \emph{Resonant Rite}, or the Keeper spends an SB tied to psionic/occult activity.
On fill: apply the last-Patron \textbf{benefit \& drawback} (and echo to followers/retainers); then reset.

\paragraph*{Resonant Rites.}
Designated Low Rites may \emph{optionally} mark +1 Corruption on success for added weight. Player chooses to resonate or not.

\paragraph*{Results.}
\textbf{Success:} Rite as written. \quad
\textbf{Partial:} $-1$ step or shorter duration; mark Fatigue~1. \quad
\textbf{Failure/Interrupted:} No effect; mark Fatigue~1; GM gains +1 SB (Hearts).

\paragraph*{Push.}
Resolve now; mark Fatigue~1; mark +1 Corruption; trigger a GM \textbf{Story Beat} (Patron/Road/social fallout).

\paragraph*{Song Synergy (Tight Rulings).}
\begin{itemize}
  \item \textbf{Compatible Songs} = same Patron \emph{or} clearly similar thematic purpose.
  \item \textbf{Harmony:} Two compatible Songs: +1 Effect to both.
  \item \textbf{Counterpoint:} Opposed Songs may cancel a drawback (GM adjudicates).
  \item \textbf{Chorus:} Multiple singers amplify (+1 Effect per participant), but cap coordinated \emph{Song Weaver} style combos at \textbf{3 participants}.
\end{itemize}

\paragraph*{Repertoire (Optional, Light).}
Single \textbf{Repertoire Clock [6]} for breadth: 2~seg = base DV $-1$ (min 2); 4~seg = +1 die to Song rolls; 6~seg = one \emph{temporary} Standard Rite as Song (practice-dependent).

\paragraph*{High Cantor (Prestige, Fast Ruling).}
Standard Rites as \textbf{High Cant}: instant; +1 die to primary effect; mark +1 Corruption. Recognizably flashy; repeated uses in a scene add +1 DV to subsequent \emph{Resolve} saves (fear/charm/social pressure).

\paragraph*{Bookkeeping Cap.}
At the table, track \emph{only}: \textbf{Corruption Clock} and (optionally) the single \textbf{Repertoire Clock}. Do not add per-Song timers; use outcomes and Push to pace.

\subsection*{Inspire Chorus}
\label{subsec:inspire-chorus}

While \emph{actively singing a Song} (from the action to begin until it resolves, or while a \emph{Lingering Verse} persists), the Cantor may \textbf{invoke Inspire Chorus}:

\begin{itemize}
  \item \textbf{Effect:} All \textbf{allies within Near} (including the Cantor) \textbf{shift Position +1} for \textbf{one exchange} (e.g., Desperate$\to$Risky, Risky$\to$Controlled). Position cannot exceed \textbf{Controlled}. This does not stack with other Position-shift auras; use the best single shift.
  \item \textbf{Use:} \textbf{Once per scene} at no cost. \textbf{Additional uses in the same scene} are allowed, but each \textbf{immediately marks +1 segment on the Cantor's Corruption Clock}.
  \item \textbf{Requirements:} The performance must be perceptible to recipients (line of hearing; \textit{Silence} or similar effects suppress it).
  \item \textbf{Timing:} Declare on starting the Song or at any time before it resolves; the shift lasts until the start of the Cantor's next turn.
  \item \textbf{Notes:} Using \emph{Inspire Chorus} does not change Song DV, Action cost, or outcomes. It respects \emph{Bookkeeping Light}: no new clocks are created.
\end{itemize}

\subsection*{Cantors as Cult Leaders (Chorus-Founders)}
Cantors gather crowds—and crowds gather debts. The Song’s Corruption stains the air, and listeners answer with vows, tithes, and favors. Many Cantors drift into leadership not by decree but by \emph{obligation}: their audience becomes a \textit{chorus} that expects guidance, protection, and more songs. In practice, the Cantor’s rising \textbf{Corruption} is mirrored by the flock’s growing \textbf{Obligation} to the Cantor (and the Patron behind the music).

\begin{fatebox}[Chorus Cult --- Quick Rule]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Trigger} & After a public Song using \emph{Inspire Chorus} or a \textbf{Resonant Rite} before 10+ witnesses, the Cantor may found or deepen a \textit{Chorus} (cult). \\[2pt]
\textbf{Cost} & Immediately convert \textbf{+1 Corruption segment} into \textbf{+1 Obligation} (to the Patron or the Chorus, GM’s call). \\[2pt]
\textbf{Benefit} & Gain a \textbf{Minor Follower (Chorus)}: once/scene (if present or reachable), \textit{+1 die} to Performance/Sway \emph{or} establish a rumor/cover within the community. Scale \(\approx\) Cantor’s \textbf{Presence}. \\[2pt]
\textbf{Maintenance} & Each scene/session you leverage the Chorus, mark \textbf{+1 Obligation}. If neglected, start \textbf{Devotion Sours [4]}; on fill, the Chorus fractures into a Complication (rival sect, scandal, or betrayed devotee). \\[2pt]
\textbf{Safety Valve} & During Downtime, a \textit{Vigil} (public service, free performance, or restitution) clears \textbf{1 Obligation} to the Chorus and resets \textbf{Devotion Sours} by 1. \\
\bottomrule
\end{tabularx}
\end{fatebox}


\section*{Free Casting (TAGS System)}
Some casters do not prepare rote rites. They shape raw forces through shared arcane grammar known as \textbf{TAGS}. A spell is constructed at the table using a short phrase of TAGS. You only need the fiction, the TAG selection, and a casting roll.

\subsection*{Spell Structure}
\textbf{Intent} + \textbf{Target} + \textbf{Tags} = effect.

Example formula:
\begin{quote}
``I unleash Burning • Area • Force against the marauders.''
\end{quote}

The GM sets a Difficulty Value (DV) based on TAG complexity and danger.

\subsection*{Base Difficulty Value (DV)}
Start at DV 1 and add +1 for each TAG used.

\begin{center}
\textbf{DV = 1 + number of TAGS}
\end{center}

Adding powerful or perilous TAGS (Teleportation, Transformation, Dominate) adds +2 instead.

Mastery, focus, or appropriate tools may lower DV by 1.

\subsection*{Casting Roll}
Roll \textbf{Wits + Arcana} (or Ritual, Channeling, etc.).  
Success = spell goes off.  
Failure or 1 = Backlash (see below).

\subsection*{Backlash}
Whenever a Free Caster fails—or pushes power beyond safety—the magic pushes back. Choose one:
\begin{itemize}
\item Harm 2 (Arcane)
\item +2 Fatigue
\item Corruption +1
\item Catastrophic side effect (GM describes)
\end{itemize}

If the spell included a ``Dangerous'' TAG, Backlash triggers on \emph{mixed} results as well.

\newpage

\section*{TAG Library}
Pick 1–3 for minor spells.  
Pick 4–6 for heavy magic (very dangerous).  
More than 6 is suicidal.

\subsection*{Elemental TAGS}
\begin{itemize}[leftmargin=*]
\item \textbf{Burning}: flame, heat, combustion.
\item \textbf{Freezing}: ice, slowing, brittle shatter.
\item \textbf{Storm}: lightning, crackling arcs, thunder shock.
\item \textbf{Stone}: walls, spikes, tremors, armor.
\item \textbf{Wave}: crushing water, currents, pressure.
\item \textbf{Wind}: levitate, gusts, deflection.
\end{itemize}

\subsection*{Force TAGS}
\begin{itemize}[leftmargin=*]
\item \textbf{Force}: pure kinetic power, shields, blasts.
\item \textbf{Area}: cone, circle, corridor, zone.
\item \textbf{Strike}: single target precision.
\item \textbf{Wall}: barrier or blockade.
\item \textbf{Bind}: restrain, hold, suspend.
\item \textbf{Dispel}: suppress magic, unravel effects.
\end{itemize}

\subsection*{Mind \& Veil TAGS}
\begin{itemize}[leftmargin=*]
\item \textbf{Veil}: conceal, blur, illusion, silence.
\item \textbf{Scry}: reveal hidden, see distance, read traces.
\item \textbf{Memory}: erase, alter, restore.
\item \textbf{Command}: compel short action.
\item \textbf{Fear}: panic, flee, break morale.
\end{itemize}

\subsection*{Life \& Body TAGS}
\begin{itemize}[leftmargin=*]
\item \textbf{Mend}: close wounds, restore flesh, reduce Harm 1.
\item \textbf{Purify}: remove poison, corruption, disease.
\item \textbf{Strengthen}: enhance body, armor, senses.
\item \textbf{Waken}: counter sleep, paralysis, stun.
\item \textbf{Beast}: speak with or influence animals.
\end{itemize}

\subsection*{Space \& Motion TAGS (Always +2 DV Each)}
\begin{itemize}[leftmargin=*]
\item \textbf{Leap}: jump far, blink across short space.
\item \textbf{Fold}: short-range teleport, vanish–reappear.
\item \textbf{Gate}: long distance passage, open/close path.
\item \textbf{Gravity}: crush, lift, suspend, walk skyward.
\end{itemize}

\subsection*{Creation \& Transformation TAGS (Always +2 DV Each)}
\begin{itemize}[leftmargin=*]
\item \textbf{Create}: manifest matter briefly.
\item \textbf{Summon}: call a being or construct.
\item \textbf{Transmute}: turn one thing into another.
\item \textbf{Animate}: make objects act with intent.
\end{itemize}