% -\textit{- coding: utf-8 -}-
% Chapter: Combat and Conflict
% Comprehensive Resource Guide for Fate's Edge

\chapter{Combat and Conflict}\label{chap:combat_conflict}
\index{Combat}\index{Conflict}

\section{Core Philosophy}\label{sec:combat_philosophy}
\index{Combat!Core philosophy}

Combat in \textbf{Fate's Edge} uses the same consequence‑forward mechanics as all other challenges. Every action can produce triumph and complication, and outcomes cascade through Story Beats (SB), timers, and position changes. The focus is narrative positioning, tactical choice, and meaningful risk---not granular simulation.\index{Story Beats}\index{Timers}\index{Position}

\section{Combat Structure}\label{sec:combat_structure}
\index{Combat!Structure}

\begin{tcolorbox}[title={Tracking NPC Mechanics},colback=red!5!white,colframe=red!75!black,sharp corners]
Not every meter needs to be tracked for NPCs.
\begin{itemize}
    \item \textbf{Spotlight First:} NPCs only carry Obligation, Corruption, or similar mechanics if these traits matter to the current story.
    \item \textbf{Skip the Bookkeeping:} Do not track every enemy's resource pool. If it's not driving narrative tension, it can be abstracted away.
    \item \textbf{Focus on Impact:} Apply NPC Obligation or Corruption only when it changes how the party experiences them --- e.g., a Patron visibly twisting a rival's fate, or a recurring villain consumed by corruption.
    \item \textbf{Player‑Facing First:} Keep full mechanics for PCs, since their journey is the story's core.
\end{itemize}
This principle keeps GM effort focused where it matters: driving story beats and consequences, not filling ledgers.
\end{tcolorbox}

\subsection{Rounds and Turns}\label{subsec:rounds_turns}
\index{Combat!Rounds}\index{Combat!Turns}
\begin{itemize}
    \item \textbf{Rounds:} A few seconds of simultaneous action.
    \item \textbf{Turns:} One significant action per participant.
    \item \textbf{Scenes:} A battle is usually one scene unless the fiction splits it.
    \item \textbf{Initiative:} Act in the order that makes sense fictionally; the GM adjudicates flow based on Position and established threats.\index{Combat!Initiative}
\end{itemize}

\subsection{Action Declaration}\label{subsec:action_declaration}
\index{Combat!Action declaration}
\begin{enumerate}
    \item \textbf{Approach:} Describe intent and method.
    \item \textbf{Position:} GM sets \textit{Dominant}, \textit{Controlled}, or \textit{Desperate}.\index{Position}
    \item \textbf{Pool:} Build dice (Attribute + Skill + situational modifiers).
    \item \textbf{Roll \& Resolve:} Compare to DV and the Outcome Matrix; any \textbf{1}s generate SB.\index{Difficulty Value (DV)}\index{Story Beats}
\end{enumerate}

\section{Position States}\label{sec:position_states}
\index{Combat!Position states}
\begin{description}
    \item[\indexterm{Dominant}] Advantage (cover, flank, surprise, elevation). Failure leaves options; consequences are minor.\index{Combat!Dominant}
    \item[\indexterm{Controlled}] Even footing; typical exchange of blows. Failure has teeth; moderate consequences.\index{Combat!Controlled}
    \item[\indexterm{Desperate}] Bad footing, outnumbered, wounded. Failure is severe; success may unlock extra narrative rewards.\index{Combat!Desperate}
\end{description}

\section{Range Bands and Movement}\label{sec:range_bands}
\index{Combat!Range bands}\index{Combat!Movement}

\subsection{Range Band Definitions}\label{subsec:range_definitions}
\index{Combat!Range bands!Definitions}
\begin{description}
    \item[\indexterm{Close}] Arm's length; grappling; only melee is practical.\index{Combat!Close range}
    \item[\indexterm{Near}] Same room / zone; most actions occur here.\index{Combat!Near range}
    \item[\indexterm{Far}] Same site but distant; requires movement to engage.\index{Combat!Far range}
    \item[\indexterm{Absent}] Off‑screen / another area; requires significant effort / time.\index{Combat!Absent range}
\end{description}

\subsection{Movement Rules}\label{subsec:movement_rules}
\index{Combat!Movement rules}
\begin{itemize}
    \item \textbf{1 Move:} Shift one band (Close $\leftrightarrow$ Near or Near $\leftrightarrow$ Far).
    \item \textbf{Dash (action):} Shift two bands (Close $\rightarrow$ Far or Far $\rightarrow$ Close).
    \item \textbf{Engage:} Entering Close from Near usually costs a Move.
    \item \textbf{Disengage:} Leaving Close may require a test if threatened.
\end{itemize}

\section{Combat Actions}\label{sec:combat_actions}
\index{Combat!Actions}

\subsection{Standard Actions}\label{subsec:standard_actions}
\index{Combat!Standard actions}
\begin{description}
    \item[\indexterm{Attack}] Strike with the appropriate Skill (e.g., Melee or Ranged).\index{Combat!Attack}
    \item[\indexterm{Defend}] Active defence against incoming harm (parry, block, roll aside).\index{Combat!Defend}
    \item[\indexterm{Maneuver}] Change Position, create advantage, or set up an ally.\index{Combat!Maneuver}
    \item[\indexterm{Use Object}] Doors, levers, lanterns, terrain, or gear.\index{Combat!Use object}
    \item[\indexterm{Cast Spell}] Perform magical actions (see \autoref{chap:magic_system}).\index{Combat!Cast spell}
\end{description}

\subsection{Special Actions}\label{subsec:special_actions}
\index{Combat!Special actions}
\begin{description}
    \item[\indexterm{Aid}] Provide assistance to another's action (costs as per Assist rules).\index{Combat!Aid}
    \item[\indexterm{Ready}] Prepare an action with a clear trigger.\index{Combat!Ready}
    \item[\indexterm{Withdraw}] Attempt to disengage safely.\index{Combat!Withdraw}
    \item[\indexterm{Sprint}] Spend your action to Dash (two‑band shift).\index{Combat!Sprint}
\end{description}

\begin{tcolorbox}[title={Quick Rule: Defend},colback=blue!5!white,colframe=blue!75!black]
Spend your action to brace for danger.

Until next turn:
\begin{itemize}
    \item Roll to resist any attack or effect.
    \item Your Position improves by one step.
\end{itemize}
Success negates the hit. Partial reduces it. Miss means you take it---\emph{but you learn from it}.
\end{tcolorbox}

\section{Ranged Combat}\label{sec:ranged}
\index{Combat!Ranged}

\subsection{Firing Profiles}\label{subsec:firing_profiles}
\begin{description}
    \item[\indexterm{Fast}] (knives, slings, shortbows) --- Move \emph{and} Shoot. Reload is incidental. Controlled at Close while Engaged.
    \item[\indexterm{Standard}] (longbows, crossbows) --- \emph{Move or Shoot}. Reload takes part of the turn. May \emph{Aim} for +1 die or +1 Effect.
    \item[\indexterm{Slow}] (greatbows, arbalests, siege) --- Requires Set / Brace. Reload takes a full turn. Cannot Move and Shoot in the same round.
\end{description}

\subsection{Options}\label{subsec:ranged_options}
\begin{itemize}
    \item \textbf{Aim:} Spend a turn to gain +1 die or +1 Effect on the next shot (lost if you Move or are disrupted).
    \item \textbf{Volley:} Spend extra ammunition for +1 die (max +2).
    \item \textbf{Suppress:} Target a zone; foes act at --1 die or Limited Effect until they shift cover.
    \item \textbf{Overwatch:} Hold action; take a Controlled shot when a foe crosses your line.
\end{itemize}

\subsection{Cover \& Close}\label{subsec:cover_close}
\begin{itemize}
    \item \textbf{Light cover:} +1 DV or --1 die to attacker.
    \item \textbf{Heavy cover:} +2 DV or --2 dice.
    \item \textbf{Obscured (fog, night):} Worsen Position by one step.
    \item \textbf{Shooting while Engaged at Close:} Controlled (Fast) or Desperate (Standard / Slow).
\end{itemize}

\section{Health, Fatigue, and Harm}\label{sec:health_fatigue_harm}
\index{Fatigue}\index{Harm}

\subsection{Tracks and Caps}
\begin{itemize}
    \item \textbf{Fatigue Track:} boxes equal to \textbf{Body}.
    \item \textbf{Harm Levels:} Harm 1, Harm 2, Harm 3.
\end{itemize}

\subsection{Fatigue $\rightarrow$ Harm Conversion}
When your Fatigue Track \emph{fills}:
\begin{enumerate}
    \item Increase your \textbf{Harm} by one level (0 $\rightarrow$ Harm 1, 1 $\rightarrow$ 2, 2 $\rightarrow$ 3).
    \item \textbf{Clear all Fatigue}.
\end{enumerate}
This can occur multiple times in a scene.

\subsection{Taking Fatigue}
Mark Fatigue for strain, exertion, travel, magic costs, or Obligation overflow. If an instance of Fatigue would exceed remaining boxes, only fill to trigger the conversion; any excess is ignored after Harm increases and the track clears.

\subsection{Recovering Fatigue}
\begin{itemize}
    \item \textbf{Short Rest} (quiet watch, food/water): remove \textbf{2 Fatigue}.
    \item \textbf{Full Night:} remove \textbf{all Fatigue}.
\end{itemize}
Fatigue recovery does not remove Harm (use medical / ritual rules).

\subsection{Harm and Consequences}\label{subsec:harm_consequences}
\index{Combat!Harm}\index{Combat!Consequences}
\begin{description}
    \item[\indexterm{Harm 1}] --1 die to related actions.
    \item[\indexterm{Harm 2}] --1 die to most actions until treated.
    \item[\indexterm{Harm 3}] Incapacitated or dying.
\end{description}
\textbf{Resist:} Roll a relevant Attribute vs. DV 3 (1s still generate SB) to reduce harm by one level.

\section{Teamwork in Combat}\label{sec:teamwork}
\index{Combat!Teamwork}

\subsection{Assistance}\label{subsec:assistance}
\index{Assists!Maximum}
\begin{itemize}
    \item \textbf{Cost:} 1 Boon.\index{Boons}
    \item \textbf{Effect:} +1 die to the assisted roll (max +3 from all assists unless a Talent says otherwise).\index{Assists!Cap}
\end{itemize}

\subsection{Setup Actions}\label{subsec:setup_actions}
\begin{itemize}
    \item Create cover, draw fire, threaten flanks, or break formation.
    \item On success, grant +1 Position \emph{or} step up Effect to an ally's next action.
\end{itemize}

\subsection{Protection}\label{subsec:protection}
\begin{itemize}
    \item Interpose to take harm intended for an ally.
    \item You must be in plausible Position; resolve as a defence or resist.
\end{itemize}

\section{Tactical Timers}\label{sec:tactical_timers}
\index{Combat!Tactical timers}
Use timers to track persistent pressures and battlefield states.

\begin{longtable}{p{4cm} p{9cm}}
\caption{Tactical Timer Examples}\label{tab:tactical_timers}\\
\toprule
\textbf{Timer Name} & \textbf{Description} \\
\midrule
Mob Overwhelm [6] & Numbers begin to swamp the PCs. \\
Fatigue Spiral [4] & Exhaustion degrades performance. \\
Morale Collapse [6] & A side is on the brink of routing. \\
Environmental Collapse [8] & Fire, flood, or structure failure escalates. \\
Reinforcement Arrival [4] & Additional foes or allies appear. \\
\bottomrule
\end{longtable}

\section{Position Dynamics}\label{sec:position_dynamics}
\index{Combat!Position dynamics}

\subsection{GM‑Initiated Shifts}\label{subsec:gm_shifts}
\begin{itemize}
    \item \textbf{Spend 1 SB:} Worsen a character's Position by one step.\index{Story Beats}
    \item \textbf{Narrative Events:} Reinforcements, collapsing cover, weather, smoke, crowding.
\end{itemize}

\subsection{Player‑Initiated Shifts}\label{subsec:player_shifts}
\begin{itemize}
    \item \textbf{Spend 1 Boon:} Improve Position by one step for the current action.\index{Boons}
    \item \textbf{Maneuvers / Assets:} Flank, gain elevation, trigger tools / terrain / followers.
\end{itemize}

\section{Magic in Combat}\label{sec:magic_combat}
\index{Combat!Magic integration}

\subsection{Casting Actions}\label{subsec:casting_actions}
\begin{description}
    \item[\indexterm{Standard Casting}] \textit{Channel} then \textit{Weave}: 1 action each, in order.
    \item[\indexterm{Rushed Casting}] Combine phases at \textit{Desperate} with harsher consequences.
    \item[\indexterm{Rites Invocation}] 1 action; may \textit{Push} for +1 Obligation.\index{Obligation}
    \item[\indexterm{Invoker Rituals}] Usually too slow; \textit{Crack the Seal} for instant effect at cost.
\end{description}

\subsection{Combat Spell Effects}\label{subsec:spell_effects}
\begin{itemize}
    \item Shift Position for multiple combatants; advance timers.
    \item Spawn hazards (smoke, grease, quake) or advantages (light, ward, barrier).
    \item Grant offensive / defensive edges; Backlash may add Fatigue / Harm or SB.\index{Backlash}
\end{itemize}

\section{Social Conflict}\label{sec:social_conflict}
\index{Social conflict}

\subsection{Social Skills}\label{subsec:social_skills}
\begin{description}
    \item[\indexterm{Sway}] Persuasion, negotiation, formal discourse.
    \item[\indexterm{Deception}] Lies, misdirection, manipulation.
    \item[\indexterm{Performance}] Oratory, entertainment, emotional appeal.
    \item[\indexterm{Insight}] Reading people, spotting tells, motives.
    \item[\indexterm{Command}] Leadership, intimidation, asserting authority.
\end{description}

\subsection{Social Position \& Consequences}\label{subsec:social_position}
\begin{description}
    \item[\indexterm{Dominant}] You hold leverage, information, or status.
    \item[\indexterm{Controlled}] Even footing; standard negotiation.
    \item[\indexterm{Desperate}] You lack leverage; they hold the cards.
\end{description}
SB may manifest as rumours / scandal, strained alliances, owed concessions, lost access, or broken relationships.

\subsection{Running Social Timers (GM Guidance)}\label{subsec:social_timers}
\index{Social conflict!Timers}\index{Persuasion}

\paragraph{Framing the Ask.}
State a concrete outcome ("grant passage tonight," "drop the bounty," "fund our expedition"). If it is strictly binary and low‑stakes, use one roll. Otherwise, build a timer.

\paragraph{Set Two Timers.}
\begin{itemize}
    \item \textbf{Persuasion} (4/6/8): your progress.
    \item \textbf{Opposition} (4/6): their resistance (ego, risk, duty, rival's whisper).
\end{itemize}
Name them (e.g., \textsc{Council Swayed}, \textsc{Captain's Doubt}) so the table sees the story move.

\paragraph{Position $\rightarrow$ DV.}
Use setting, leverage, and stakes:
\begin{itemize}
    \item \textbf{Dominant (DV 2):} private audience, proof in hand, shared values.
    \item \textbf{Controlled (DV 3):} time pressure, partial access, mixed reception.
    \item \textbf{Desperate (DV 4--5+):} public scrutiny, scandal risk, hostile crowd.
\end{itemize}

\paragraph{Distinct Approaches (examples).}
\begin{itemize}
    \item \textbf{Wits + Sway:} reframe incentives; offer face‑saving out.
    \item \textbf{Wits + Lore:} cite precedent, produce documents or testimony.
    \item \textbf{Body + Presence:} command presence, ritual authority, oath.
    \item \textbf{Wits + Tinker:} demonstrate a device / proof‑of‑concept on the spot.
    \item \textbf{Bonds / Boons:} an ally vouches (assist) or gifts a Boon; NPC bonds can reduce DV by 1 if genuinely invoked.
\end{itemize}

\paragraph{Outcomes Palette.}
\begin{itemize}
    \item \textbf{Strong Hit:} Tick Persuasion +2; optionally bank a \emph{Concession} (waive a future cost).
    \item \textbf{Mixed:} Tick +1 and choose a cost: start a small \textsc{Rebuttal 1/4}, owe a minor favor, or the GM spends 1 SB to introduce a new stakeholder.
    \item \textbf{Miss:} No progress. GM may (a) reduce Persuasion --1, (b) advance Opposition +1--2, or (c) worsen Position one step.
\end{itemize}

\paragraph{When Opposition Fills First.}
The target hardens or reframes the negotiation: narrow the ask, accept a condition, or escalate proof (pay a Boon, reveal evidence, bring a witness) to keep going.

\paragraph{Scaling \& Dials.}
\begin{itemize}
    \item \textbf{Crowd Scenes:} Add \textsc{Hecklers} / \textsc{Applause} 1/4 that swing Position when they fill.
    \item \textbf{Stake Weight:} For life‑or‑death asks, require a \emph{Concession} on success (named cost paid during wrap‑up).
    \item \textbf{Truth vs. Bluff:} Real proof improves Position; blatant lies risk a hidden \textsc{Caught Out} 1/4 that explodes later.
    \item \textbf{Patron Colour:} Invoking a Patron symbol can shift Position if on‑theme---or add +1 DV if it antagonises the audience's loyalties.
\end{itemize}

\paragraph{End States.}
On fill, summarise the agreement and record any \emph{Concessions}, debts, or timers that carry forward. If not filled by scene end, bank current ticks and reopen later if fiction supports it.

\section{Mass Combat}\label{sec:mass_combat}
\index{Combat!Mass combat}
Treat armies as high‑Cap followers with domain tags and timers.

\subsection{Army as Followers}\label{subsec:army_followers}
\begin{itemize}
    \item \textbf{Cost:} Cap\(^{2}\) XP to raise / maintain.\index{Followers!Cost}
    \item \textbf{Types:} Infantry, cavalry, archers, engineers, fleets.
    \item \textbf{Capabilities:} Provide large assist dice to war‑scale actions (still capped unless a rule overrides).\index{Assists!Maximum}
    \item \textbf{Risks:} Supply, morale, command / control, terrain.
\end{itemize}

\subsection{War Timers}\label{subsec:war_timers}
\begin{description}
    \item[\indexterm{Supply Lines [8]}] Logistics and provisioning.
    \item[\indexterm{Army Morale [6]}] Cohesion and willingness to fight.
    \item[\indexterm{Strategic Position [8]}] Passes, ports, bridges.
    \item[\indexterm{Alliance Stability [6]}] Political support and coalition strain.
\end{description}

\section{Environmental Combat}\label{sec:environmental_combat}
\index{Combat!Environmental}

\subsection{Environmental Hazards}\label{subsec:environmental_hazards}
\begin{description}
    \item[\indexterm{Fire [6]}] Spreading flames limit movement and cause harm.
    \item[\indexterm{Flood [8]}] Rising water creates difficult terrain and drowning risk.
    \item[\indexterm{Collapse [6]}] Falling debris, blocked routes.
    \item[\indexterm{Weather [4]}] Storms, fog, glare reduce visibility / accuracy.
\end{description}

\subsection{Terrain Effects}\label{subsec:terrain_effects}
\begin{itemize}
    \item \textbf{Choke Points:} Favor defenders, constrain numbers.
    \item \textbf{Elevation:} Bonuses to ranged / oversight, harder to assault.
    \item \textbf{Cover:} Improves Position and reduces consequence severity.
    \item \textbf{Difficult Terrain:} Consumes movement; may worsen Position.
\end{itemize}

\section{Miniatures and Tactical Layer}\label{sec:miniatures}
\index{Combat!Miniatures}\index{Tactical layer}

\textit{Design Note:} The core loop handles all conflict through narrative positioning. However, the GM may wish to consider the Tactical Layer as a visual aid when the fiction demands high precision (e.g., tight corridors, complex terrain, or large‑scale skirmishes). This is a modular choice; the core narrative loop remains the primary engine regardless of whether miniatures are present.

\subsection{Core Concepts}
Works on square or hex grids; declare type at setup. Units have base sizes (Small--Huge) and facing. Actions per turn: Move then Act (or vice versa). All checks use normal SRD roll + DV.

\subsection{Turn Structure}
\begin{enumerate}[nosep]
    \item \textbf{Start:} Resolve ongoing effects.
    \item \textbf{Move:} Up to Speed (typically \(5 + \text{Body}\)); obey ZoC.
    \item \textbf{Act:} Attack, cast, assist, shove, guard, etc.
    \item \textbf{End:} Resolve reactions and end‑of‑turn effects.
\end{enumerate}

\subsection{Zones of Control (ZoC)}\label{subsec:zoc}
\begin{itemize}[nosep]
    \item \textbf{Size:} 4 orthogonals (squares) or 6 adjacents (hexes). Large / Huge project ZoC from edges; Reach may extend ZoC by +1 ring.
    \item \textbf{Entry:} Entering an enemy's ZoC ends your movement (you are \textbf{Engaged}).
    \item \textbf{Passage:} You cannot move \textit{through} an enemy's ZoC unless you \textbf{Push‑through} (below) or \textbf{Disengage} (Action; DV 4).
\end{itemize}

\subsubsection*{Push‑Through}
Force movement through an enemy's ZoC without Disengaging:
\begin{itemize}[nosep]
    \item \textbf{Test:} Body + Athletics vs. DV \(= 2 + \text{enemy Tier}\) (or Cap for followers / minions).
    \item \textbf{Cost (choose one):}
        \begin{itemize}[nosep,label=$\bullet$]
            \item Drop Position by one step (min. Desperate) on next action this turn, \textbf{or}
            \item Suffer --1 die on next roll, \textbf{or}
            \item Grant enemy immediate opportunity attack at +1 die.
        \end{itemize}
    \item \textbf{Fail:} Stop in their ZoC; movement ends.
    \item \textbf{Tier IV+:} Spend 1 Boon to ignore ZoC for one Move (treat as narrative space).
\end{itemize}

\subsection{Facing and Flanking}\label{subsec:facing_flanking}
Choose facing at end of movement.
\begin{itemize}[nosep]
    \item \textbf{Flank} (opposite arcs): +1 die to attack.
    \item \textbf{Rear} (behind facing): +1 die and +1 Effect.
    \item \textbf{Assist (mini):} +1 Effect (not dice); max 1 helper.
\end{itemize}

\subsection{Special Actions (Miniatures)}\label{subsec:special_actions_mini}
\begin{itemize}[nosep]
    \item \textbf{Guard:} Ready strike when enemy leaves your ZoC.
    \item \textbf{Dash:} +2 movement this turn.
    \item \textbf{Brace:} Resist Shoves; extend ZoC for opportunity attacks only.
    \item \textbf{Shove / Tackle:} Body + Athletics vs. DV 4--6 to push target 1 hex or knock prone.
\end{itemize}

\subsection{Magic Integration (Miniatures)}\label{subsec:magic_mini}
Magic uses \textbf{[TAGS]} (e.g., \texttt{[WARD]}, \texttt{[AREA]}). Casting while Engaged worsens Position unless \texttt{[INSTANT]} or aided by Talent. Rituals require clear space; disrupted rituals fail or require Spirit + Resolve to maintain.

\begin{tcolorbox}[title={Miniatures Quick Reference},colback=white!98!gray,colframe=black!50!gray,boxrule=0.4pt,fontupper=\small]
\textbf{ZoC:} Enter / leave adjacent hex $\rightarrow$ provokes 1 Reaction (Free Strike \emph{or} Shove 1 hex). Max 1 Reaction / round.

\textbf{Position:} Set by fiction (flank / rear improve; cover improves vs. ranged).

\textbf{DV Mods:} Elevation +1; Difficult terrain $2\times$ MP; Disengage $=4$.

\textbf{Tier Scale:} At Tier IV+, may spend 1 Boon to ignore ZoC for one Move.

\textbf{Crit:} Bump Position one step (max Dominant $\rightarrow$ Push 1 hex or +1 Success).

\textbf{Tags:} Max 2 active tags / unit. \texttt{[WARD]} $= -1$ die; attacker may take 2 Fatigue to ignore.
\end{tcolorbox}

\section{Quick Reference}\label{sec:combat_quick_ref}
\index{Combat!Quick reference}

\subsection{Position Effects}\label{subsec:position_quick_ref}
\index{Combat!Quick reference!Position}
\begin{longtable}{lll}
\toprule
\textbf{Position} & \textbf{Typical Edge} & \textbf{Consequence Severity} \\
\midrule
Dominant   & Better options, easier withdraw & Minor \\
Controlled & Standard options               & Moderate \\
Desperate  & High reward potential          & Severe \\
\bottomrule
\end{longtable}

\subsection{Harm Quick Reference}\label{subsec:harm_quick_ref}
\index{Combat!Quick reference!Harm}
\begin{longtable}{llll}
\toprule
\textbf{Harm} & \textbf{Penalty} & \textbf{Typical SB Grant} & \textbf{Recovery} \\
\midrule
Harm 1 & --1 die (related)      & 0--1 (GM option) & Rest / basic care \\
Harm 2 & --1 die (most)         & 0--1 (GM option) & Treatment \\
Harm 3 & Incapacitated         & 1--2 (GM option) & Major intervention \\
\bottomrule
\end{longtable}

\subsection{Common Action DVs}\label{subsec:action_dvs}
\index{Combat!Quick reference!DVs}
\begin{longtable}{lll}
\toprule
\textbf{Action Type} & \textbf{Typical DV} & \textbf{Notes} \\
\midrule
Basic Attack    & 2 & Standard melee / ranged in even footing \\
Maneuver        & 2--3 & Create advantage, change Position \\
Active Defence  & 3 & Parry, block, evade under pressure \\
Complex Action  & 4 & Multi‑target or strong field control \\
High‑Risk       & 5+ & Desperate gambit, extreme precision \\
\bottomrule
\end{longtable}

\section{Combat Examples}\label{sec:combat_examples}
\index{Combat!Examples}

\subsection{Melee Combat Example}\label{subsec:melee_example}
\index{Combat!Examples!Melee}
Mateo strikes a cultist (\textit{Controlled}, DV 2):
\begin{itemize}
    \item 5d10 $\rightarrow$ 9, 7, 5, 2, 1 $\Rightarrow$ 3 successes, 1 SB.
    \item \textbf{Success with SB:} The cultist falls.
    \item GM spends 1 SB: \emph{Blood spatters the ritual circle; the summoning timer advances 1.}
\end{itemize}

\subsection{Ranged Combat Example}\label{subsec:ranged_example}
\index{Combat!Examples!Ranged}
Lyra fires a longbow at a distant archer (\textit{Desperate}, DV 3):
\begin{itemize}
    \item 4d10 $\rightarrow$ 10, 6, 3, 1 $\Rightarrow$ 2 successes, 1 SB.
    \item \textbf{Partial:} The archer is hit but dives for cover.
    \item GM offers choice: shift Lyra to \textit{Desperate} (stay exposed) \emph{or} take Harm 1 from return fire.
\end{itemize}

\subsection{Magic Combat Example}\label{subsec:magic_example}
\index{Combat!Examples!Magic}
Theron raises a defensive ward (\textit{Desperate}, DV 3):
\begin{itemize}
    \item \textbf{Channel:} 2 successes, 0 SB.
    \item \textbf{Weave:} 3 successes, 2 SB.
    \item \textbf{Success with SB:} Ward holds; GM spends 2 SB to start \emph{Ward Strain [4]}.
\end{itemize}

\section{Tags \& States (Combat)}\label{sec:tags_states_ref}
\begin{description}
    \item[\texttt{[HEALED]}] Remove \textbf{all Fatigue}. (Does not remove Harm.)
    \item[\texttt{[RALLIED]}] Remove \textbf{2 Fatigue}; +1 die on your next action this scene.
    \item[\texttt{[FORTIFIED]}] Until scene end, the first incoming \textbf{Harm 1} becomes \textbf{2 Fatigue}.
    \item[\texttt{[STABILIZED]}] End bleed / burn / poison; does not remove Harm.
    \item[\texttt{[MENDED]}] Remove Harm per your recovery procedure.
    \item[\texttt{[REVIVED]}] Stand a fallen ally; follow Harm 3 / KO rules.
\end{description}

\section{Melee Combat Talents}\label{sec:melee_talents}
\index{Talents!Combat}\index{Combat!Talents}

\subsection{Minor Talents}

\subsubsection{Defensive Survival (3 XP)}\label{talent:defensive_survival}
\textit{Prerequisite:} Melee 2+.
\textbf{Effect:} +1 to defense DV while engaged in melee. Once per scene, convert first Harm 1 from melee to Fatigue.
\textit{Narrative:} Years of combat teaching you to read attacks and flow with them.

\subsubsection{Tactical Movement (4 XP)}\label{talent:tactical_movement}
\textit{Prerequisite:} Athletics 2+.
\textbf{Effect:} Move within engagement zone as Move action (instead of full action). Once per scene, disengage from Close as Move action.
\textit{Narrative:} Footwork and positioning that keeps you alive in the press.

\subsubsection{Conditioning (4 XP)}\label{talent:conditioning}
\textit{Prerequisite:} Body 3+.
\textbf{Effect:} Body attribute counts as +1 for Fatigue track calculations. +1 die to resist Fatigue overflow effects.
\textit{Narrative:} Physical conditioning that lets you endure punishment.

\subsubsection{Weapon Master (5 XP)}\label{talent:weapon_master}
\textit{Prerequisite:} Melee 2+.
\textbf{Effect:} +2 dice (instead of +1) with chosen weapon category. Once per scene, +1 Effect with signature weapon.
\textit{Narrative:} Mastery of specific weapons that makes them extensions of yourself.

\subsection{Major Talents}

\subsubsection{Flurry Strike (7 XP)}\label{talent:flurry_strike}
\textit{Prerequisites:} Melee 3+, Body 3+.
\textbf{Effect:} When engaged with multiple opponents, make 2 attacks as one action. Each attack at --1 die.
\textit{Narrative:} Training that lets you fight multiple enemies simultaneously.

\subsubsection{Duelist's Edge (8 XP)}\label{talent:duelist_edge}
\textit{Prerequisites:} Melee 3+, Wits 3+.
\textbf{Effect:} When engaged with single opponent: +1 die to all melee rolls. Once per scene, ignore first Harm 1 or 2 from that opponent.
\textit{Narrative:} Psychological and tactical dominance in one‑on‑one combat.

\subsubsection{Battlefield Mastery (8 XP)}\label{talent:battlefield_mastery}
\textit{Prerequisites:} Melee 4+, Wits 4+, Command 2+.
\textbf{Effect:} Once per scene, when engaged with 3+ opponents, declare "Battlefield Mastery." For next 3 exchanges:
\begin{itemize}[nosep]
    \item All melee attacks gain +1 Effect.
    \item Enemies act at --1 die due to disorientation.
    \item Your Position improves by one step.
    \item Convert one Harm 1 $\rightarrow$ Fatigue per exchange.
\end{itemize}
\textit{Narrative:} When surrounded, you enter a state of perfect combat flow where enemies become obstacles rather than threats.

\subsection{Prestige Talents}

\subsubsection{Battlefield Terror (12 XP)}\label{talent:battlefield_terror}
\textit{Prerequisites:} Melee 4+, Body 4+, Harm 2+ experience.
\textbf{Effect:} Enemies in Close range act at --1 die due to intimidation. Once per scene, convert enemy's success to partial with cost.
\textit{Narrative:} Reputation and presence that makes opponents hesitate.

\subsection{Epic Talents}

\subsubsection{Blade Dance (18 XP)}\label{talent:blade_dance}
\textit{Prerequisites:} Melee 5+, Duelist's Edge, Flurry Strike.
\textbf{Effect:} Engage and attack up to 3 targets in one action. Each attack at --1 die, but Position improves by one step.
\textit{Narrative:} Legendary skill that makes you a whirlwind of death.

\subsection{Combat Balance Notes}
These talents are designed to enhance melee viability while maintaining \textbf{Fate's Edge}'s core tension between risk and reward. Melee combat should remain \textbf{manageably deadly} -- dangerous enough to require tactical skill, but with meaningful options for skilled fighters to excel.

\paragraph{Key Principles:}
\begin{itemize}
    \item Talents enhance existing mechanics rather than replace them.
    \item Specialisation provides clear advantages for focused builds.
    \item High‑Cap opponents remain genuinely threatening.
    \item Positioning and tactical decision‑making remain crucial.
    \item Story Beat escalation continues to compound challenges.
\end{itemize}

\paragraph{Role Balance:}
Enhanced melee fighters complement rather than overshadow other roles. Ranged characters maintain mobility advantages, magic users provide battlefield control, and support characters enable team effectiveness.

\section{Monk Talents}\label{sec:monk_talents}
\index{Talents!Monk}\index{Monk}

\subsection{Core Concept}
Monks channel inner discipline into supernatural martial prowess, combining unarmed combat mastery with spiritual focus.

\subsection{Starting Talent}
\subsubsection{Disciplined Body (3 XP -- Minor)}\label{talent:disciplined_body}
\textit{Prerequisites:} Melee 1+, Body 2+.
\textbf{Benefits:}
\begin{itemize}
    \item +1 die to unarmed combat attacks.
    \item Convert 1 Harm to Fatigue once per scene.
    \item Once per scene, improve Position by one step.
\end{itemize}

\subsection{Advanced Talents}
\subsubsection{Iron Fist Way (6 XP -- Minor)}\label{talent:iron_fist_way}
\textbf{Benefits:} +1 die to unarmed attacks; strikes count as enchanted.

\subsubsection{Flowing Spirit Way (8 XP -- Major)}\label{talent:flowing_spirit_way}
\textbf{Benefits:} Convert up to 1 Harm into Fatigue per attack; +1 die against fear or charm.

\subsubsection{Perfect Timing Way (7 XP -- Major)}\label{talent:perfect_timing_way}
\textbf{Benefits:} Twice per scene, improve Position by +1 step; +1 die to reactions.

\subsubsection{Untouchable Way (12 XP -- Major)}\label{talent:untouchable_way}
\textit{Prerequisites:} Iron Fist + Flowing Spirit.
\textbf{Benefits:} +1 die to unarmed attacks; convert 2 Harm into Fatigue; cannot be grappled.

\subsubsection{Inevitable Way (15 XP -- Major)}\label{talent:inevitable_way}
\textit{Prerequisites:} Iron Fist + Perfect Timing.
\textbf{Benefits:} +2 dice to unarmed attacks; ignore 1 Armor; may counterattack when an enemy misses.

\subsubsection{Transcendent Harmony (18 XP -- Epic)}\label{talent:transcendent_harmony}
\textit{Prerequisites:} Flowing Spirit + Perfect Timing, Spirit 4+.
\textbf{Benefits:} Convert 2 Harm into Fatigue; once per session become immune to Harm; allies gain +1 defence.

\subsection{Progression Path}
Monks specialise early (6--8 XP), combine paths mid‑tier (12--15 XP), and achieve transcendence late (18 XP). Each path represents a distinct combat philosophy and playstyle.

% End of chapter