i\chapter{Appendices}
\label{chap:appendices}

%----------------------------------
% Visual setup for quick-ref tables
%----------------------------------
\definecolor{FeRow}{HTML}{F5F7FA}
\renewcommand{\arraystretch}{1.15}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{C}{>{\centering\arraybackslash}p{1.8cm}}
\newcommand{\feTableStart}{\rowcolors{2}{FeRow}{white}\small}
\newcommand{\feTableEnd}{\rowcolors{2}{}{}}
\appendix
\chapter{Quick Reference Sheets}
\label{sec:quick-reference}
\index{Appendices!Quick Reference}

\subsection{Core Mechanic Quick Reference}
\label{subsec:core-mechanic-ref}
\index{Appendices!Core Mechanic}

\begin{enumerate}
\item \textbf{Approach}: Player states intent and method (Attribute + Skill combination).\index{Core Mechanic!Approach}
\item \textbf{Execution}: Roll dice pool of d10s. Each 6+ is a success; each 1 generates 1 Story Beat.\index{Core Mechanic!Execution}
\item \textbf{Outcome}:
  \begin{itemize}
  \item \textbf{Basic} — Roll as-is; all 1s generate SB.
  \item \textbf{Detailed} — Re-roll one die showing 1.
  \item \textbf{Intricate} — Re-roll all 1s; add a positive flourish on success.
  \end{itemize}
\end{enumerate}

\subsection{Attribute and Skill Summary}
\label{subsec:attributes-skills-ref}
\index{Appendices!Attributes and Skills}

\textbf{Attributes:}
\begin{description}
\item[Body] Physical strength, endurance, athletic ability.\index{Attributes!Body}
\item[Wits] Perception, cleverness, quick thinking.\index{Attributes!Wits}
\item[Spirit] Willpower, intuition, spiritual resilience.\index{Attributes!Spirit}
\item[Presence] Charm, command, social influence.\index{Attributes!Presence}
\end{description}

\textbf{Skill Levels:}
\begin{description}
\item[0 Untrained] Rely on raw Attribute only.\index{Skills!Untrained}
\item[1 Familiar] Basic competence, occasional use.\index{Skills!Familiar}
\item[2 Skilled] Reliable training, regular practice.\index{Skills!Skilled}
\item[3 Expert] Professional mastery, notable ability.\index{Skills!Expert}
\item[4 Master] Renowned specialist, exceptional talent.\index{Skills!Master}
\item[5 Legendary] Near-mythic capability, extraordinary.\index{Skills!Legendary}
\end{description}

\subsection{Experience Point Costs}
\label{subsec:xp-costs-ref}
\index{Appendices!XP Costs}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l l @{}}
\toprule
\textbf{Improvement} & \textbf{Cost} & \textbf{Downtime} \\
\midrule
Attribute increase & New rating $\times$ 3 XP & New rating days \\
Skill increase & New level $\times$ 2 XP & New level days \\
On-Screen Follower & Cap$^2$ XP & 1--3 days \\
Minor Asset & 4 XP & 1 day \\
Standard Asset & 8 XP & 1 week \\
Major Asset & 12 XP & 1 month \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

% !TEX root = srd_main.tex
% SRD Insert: Upkeep (Condensed)

% !TEX root = srd_main.tex
% SRD Insert: Upkeep (Condensed)

\subsection*{Upkeep}\label{sec:upkeep-srd}
\index{Upkeep}\index{Downtime}\index{Followers}\index{Assets}

\textbf{Frequency.} Pay upkeep once per Downtime period.

\begin{itemize}
\item \textbf{Efficient (Higher XP, Less Time).} \emph{Cost:} Upkeep XP $= \max\big(1, \tfrac{\text{XP Acquisition Cost}}{3}\big)$. \emph{Time:} Minimal; delegation/check-in.
\item \textbf{Intensive (Lower XP, More Time).} \emph{Cost:} 1 XP. \emph{Time:} One dedicated Downtime action of significant personal attention.
\end{itemize}

\textbf{Failure.} If upkeep is not paid this Downtime, the resource degrades:
\begin{itemize}
\item \emph{Follower:} becomes \textbf{Wary} (or \textbf{Seized} if already Wary).
\item \emph{Asset:} becomes \textbf{Neglected} (or \textbf{Compromised} if already Neglected).
\end{itemize}

\paragraph{Notes.} Each follower/asset checks upkeep separately; a single Intensive action may cover a cohesive group if fiction supports it. Tie upkeep scenes to Patron themes for flavor, not discounts.

\subsection{Difficulty Value (DV) Reference}
\label{subsec:dv-reference}
\index{Appendices!DV Reference}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}>{\centering\arraybackslash}p{1.2cm} l Y @{}}
\toprule
\textbf{DV} & \textbf{Difficulty} & \textbf{Typical Situations} \\
\midrule
2 & Routine   & Clear intent, modest stakes, controlled environment \\
3 & Pressured & Time pressure, mild resistance, partial information \\
4 & Hard      & Hostile conditions, active opposition, precise timing \\
5+ & Extreme  & Multiple constraints, high precision, dramatic failure risk \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\section{Deck Usage Reference}
\label{sec:deck-reference}
\index{Appendices!Deck Reference}

\subsection{Deck Types and Meanings}
\label{subsec:deck-types}
\index{Appendices!Deck Types}

\begin{description}
\item[\textbf{Travel Decks} (regional, 52-card)] Used for journey content and location-based adventures.\index{Decks!Travel}
\begin{itemize}
\item Spade $=$ Place/Location
\item Heart $=$ Actor/Faction
\item Club $=$ Pressure/Complication
\item Diamond $=$ Leverage/Reward
\end{itemize}

\item[\textbf{Deck of Consequences} (scene drama)] Used for immediate complications and narrative twists.\index{Decks!Consequences}
\begin{itemize}
\item Hearts $=$ Social/Emotional fallout
\item Spades $=$ Harm/Escalation
\item Clubs $=$ Material cost/Resource drain
\item Diamonds $=$ Magical/Spiritual disturbance
\end{itemize}
\end{description}

\textbf{Important:} Never mix suit meanings across decks. Travel deck suits differ from Consequences deck suits.

\subsection{Deck Usage Procedure}
\label{subsec:deck-procedure}
\index{Appendices!Deck Procedure}

After a roll generating Story Beats:
\begin{enumerate}
\item \textbf{Direct Spend}: Translate SB into immediate consequences or clock ticks.
\item \textbf{Deck Draw}: Draw up to $\min(\text{SB},\,3)$ cards and synthesize a single twist.
\item Interpret cards based on suit meanings and highest rank.
\end{enumerate}

\subsection{Rank Severity Guide}
\label{subsec:rank-severity}
\index{Appendices!Rank Severity}

\begin{description}
\item[Ace--3] Minor inconvenience or flavor complication.\index{Decks!Ranks!Minor}
\item[4--6] Moderate setback with narrative impact.\index{Decks!Ranks!Moderate}
\item[7--9] Significant consequence altering the scene.\index{Decks!Ranks!Significant}
\item[10--King] Major fallout introducing new problems or lasting effects.\index{Decks!Ranks!Major}
\end{description}

\section{Magic System Quick Reference}
\label{sec:magic-reference}
\index{Appendices!Magic Reference}

\subsection{Magic Paths Comparison}
\label{subsec:magic-paths-ref}
\index{Appendices!Magic Paths}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l l l @{}}
\toprule
\textbf{Path} & \textbf{Requirements} & \textbf{Key Feature} & \textbf{Risk Type} \\
\midrule
Caster (Freeform) & Caster's Gift (2 XP) & Flexible improvisation & Backlash \\
Runekeeper (Rites) & Thiasos + Codex (6 XP) & Structured Rites & Obligation \\
Invoker (Symbols) & Patron's Symbol (4 XP) & Ritual precision & Symbol compromise \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\subsection*{DV Reference Table}
\label{tab:dv-reference}
\index{Rites!DV Table}

The following table shows the resulting DV for common Obligation Costs across Spirit scores and Rite Tiers.
DV is always calculated as $\max(\text{Obligation} - \text{Spirit}, \, \text{Tier})$.

\begin{longtable}{c|cccc|c|c|c}
\toprule
\textbf{Obligation Cost} & \textbf{Spirit 0} & \textbf{Spirit 1} & \textbf{Spirit 2} & \textbf{Spirit 3--4} & \textbf{Tier 1} & \textbf{Tier 2} & \textbf{Tier 3} \\
\midrule
\endfirsthead
\toprule
\textbf{Obligation Cost} & \textbf{Spirit 0} & \textbf{Spirit 1} & \textbf{Spirit 2} & \textbf{Spirit 3--4} & \textbf{Tier 1} & \textbf{Tier 2} & \textbf{Tier 3} \\
\midrule
\endhead
\bottomrule
\endfoot
\bottomrule
\endlastfoot
1 & 1 & 1 & 1 & 1 & 1 & 2 & 3 \\
2 & 2 & 1 & 1 & 1 & 1 & 2 & 3 \\
3 & 3 & 2 & 1 & 1 & 1 & 2 & 3 \\
4 & 4 & 3 & 2 & 1 & 1 & 2 & 3 \\
5 & 5 & 4 & 3 & 2 & 1 & 2 & 3 \\
6 & 6 & 5 & 4 & 3 & 1 & 2 & 3 \\
7 & 7 & 6 & 5 & 4 & 1 & 2 & 3 \\
\end{longtable}

\paragraph{How to Read.}
\begin{itemize}
  \item Left block: DV before applying the Tier floor.
  \item Right block: the minimum DV once Tier is considered.
  \item Example: A Rite with Obligation 4, Spirit 2, Tier 2 → Base DV = 2, but Tier floor raises it to 2.
\end{itemize}

\subsection{Casting Loop Summary}
\label{subsec:casting-loop-ref}
\index{Appendices!Casting Loop}

\begin{enumerate}
\item \textbf{Channel}: Wits + Arcana roll to gather Potential.\index{Magic!Channel}
\item \textbf{Weave}: Wits + Art roll to shape spell effect.\index{Magic!Weave}
\item \textbf{Backlash}: SB spent through thematic consequences.\index{Magic!Backlash}
\end{enumerate}

\subsection{Eight Elements of Magic}
\label{subsec:elements-ref}
\index{Appendices!Magic Elements}

\begin{description}
\item[Earth] Solidity, stability, foundation.\index{Magic!Elements!Earth}
\item[Fire] Energy, transformation, destruction.\index{Magic!Elements!Fire}
\item[Air] Movement, speed, freedom.\index{Magic!Elements!Air}
\item[Water] Fluidity, healing, adaptability.\index{Magic!Elements!Water}
\item[Fate] Destiny, inevitability, causality.\index{Magic!Elements!Fate}
\item[Life] Vitality, creation, growth.\index{Magic!Elements!Life}
\item[Luck] Chance, unpredictability, probability.\index{Magic!Elements!Luck}
\item[Death/Dreams] Endings, thresholds, subconscious.\index{Magic!Elements!Death}
\end{description}

\section{Combat and Conflict Reference}
\label{sec:combat-reference}
\index{Appendices!Combat Reference}

\subsection{Position States}
\label{subsec:position-ref}
\index{Appendices!Position States}

\begin{description}
\item[Controlled] Advantageous position, minor consequences.\index{Combat!Controlled}
\item[Risky] Standard situation, moderate consequences.\index{Combat!Risky}
\item[Desperate] Disadvantaged, severe consequences.\index{Combat!Desperate}
\end{description}

\subsection{Harm Levels and Effects}
\label{subsec:harm-ref}
\index{Appendices!Harm Reference}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l l l @{}}
\toprule
\textbf{Harm Level} & \textbf{SB Generation} & \textbf{Penalty} & \textbf{Recovery} \\
\midrule
Minor & 1 SB on next 2 rolls & $-1$ die to related actions & Rest or basic care \\
Moderate & 1 SB on next roll & $-1$ die to most actions & Medical treatment \\
Severe & 2 SB on next roll & $-2$ dice to most actions & Extended care \\
Critical & 3 SB on next roll & Incapacitated & Major intervention \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\subsection{Range Bands}
\label{subsec:range-bands-ref}
\index{Appendices!Range Bands}

\begin{description}
\item[Close] Arm's length, grappling distance.\index{Combat!Close Range}
\item[Near] Same room or immediate area.\index{Combat!Near Range}
\item[Far] Visible but not immediately reachable.\index{Combat!Far Range}
\item[Absent] Off-screen or out of current scene.\index{Combat!Absent Range}
\end{description}

\subsection{Movement Actions}
\label{subsec:movement-ref}
\index{Appendices!Movement}

\begin{itemize}
\item \textbf{1 Move}: Shift one range band (Close$\leftrightarrow$Near or Near$\leftrightarrow$Far).
\item \textbf{Dash Action}: Shift two bands in one action.
\item \textbf{Disengage}: Test to leave Close range when threatened.
\item \textbf{Sprint}: Rapid movement across the battlefield.
\end{itemize}

\section{Resource Management Reference}
\label{sec:resource-reference}
\index{Appendices!Resource Reference}

\subsection{Story Beat Economy}
\label{subsec:sb-economy-ref}
\index{Appendices!Story Beats}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}>{\centering\arraybackslash}p{1.8cm} l Y @{}}
\toprule
\textbf{SB Cost} & \textbf{Effect Scale} & \textbf{Typical Effects} \\
\midrule
1 SB & Minor pressure & Noise, trace, time loss, +1 Supply segment \\
2 SB & Moderate setback & Alarm, lose position/cover, lesser foe appears \\
3 SB & Serious trouble & Reinforcements, key gear breaks, major complication \\
4+ SB & Major turn & Trap springs, authority arrives, scene shifts dramatically \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\subsection{Boon Usage Guide}
\label{subsec:boon-usage-ref}
\index{Appendices!Boons}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}l l Y @{}}
\toprule
\textbf{Boon Cost} & \textbf{Effect} & \textbf{Limitations} \\
\midrule
1 Boon & Re-roll one die            & Once per action \\
1 Boon & Activate on-screen Asset   & Plausibility test required \\
1 Boon & Improve Position by 1 step & One step maximum per action \\
2 Boons & Convert to 1 XP           & Once per session; max 2 XP \\
Variable & Power Rites/Abilities    & As specified \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\textbf{Boon Limits:}
\begin{itemize}
\item Hold maximum of 5 Boons at any time.
\item Trim to 2 Boons at scene endings.
\item Maximum 2 Boons from failures per character per scene.
\item Conversion: 2 Boons $=$ 1 XP (max 2 XP per session).
\end{itemize}

\subsection{Supply Clock States}
\label{subsec:supply-ref}
\index{Appendices!Supply Clock}

\begin{description}
\item[\textbf{Full Supply} (0)] No penalties; well-equipped.\index{Supply!Full}
\item[\textbf{Low Supply} (2)] Minor narrative complications.\index{Supply!Low}
\item[\textbf{Dangerously Low} (3)] Each character gains 1 Fatigue.\index{Supply!Dangerously Low}
\item[\textbf{Out of Supply} (4)] Severe penalties; starvation risk.\index{Supply!Out of Supply}
\end{description}

\section{Travel and Exploration Reference}
\label{sec:travel-reference}
\index{Appendices!Travel Reference}

\subsection{Travel Clock Sizes}
\label{subsec:travel-clocks-ref}
\index{Appendices!Travel Clocks}

\begin{description}
\item[4 segments] Short, straightforward journeys.\index{Travel!Clocks!4-segment}
\item[6 segments] Standard travel legs.\index{Travel!Clocks!6-segment}
\item[8 segments] Extended or complex journeys.\index{Travel!Clocks!8-segment}
\item[10 segments] Epic or highly dangerous travel.\index{Travel!Clocks!10-segment}
\end{description}

\subsection{Card Draw Procedures}
\label{subsec:card-draw-ref}
\index{Appendices!Card Draw}

\textbf{Quick Hook (2 cards):}
\begin{itemize}
\item Draw one Spade (place) and one Heart (actor).
\item Use higher rank to set clock size.
\end{itemize}

\textbf{Full Seed (4 cards):}
\begin{itemize}
\item Draw until one card of each suit appears.
\item Spade $=$ location, Heart $=$ faction, Club $=$ pressure, Diamond $=$ leverage.
\item Highest rank sets main clock size.
\end{itemize}

\section{Character Advancement Guide}
\label{sec:advancement-reference}
\index{Appendices!Advancement}

\subsection{Reputation Tiers}
\label{subsec:reputation-tiers-ref}
\index{Appendices!Reputation Tiers}

\begin{description}
\item[Tier I -- Rookie (0--40 XP)] Local reputation; prestige locked.\index{Reputation Tiers!Rookie}
\item[Tier II -- Seasoned (41--90 XP)] Regional notice; prestige may unlock.\index{Reputation Tiers!Seasoned}
\item[Tier III -- Veteran (91--150 XP)] National influence; second follower suggested.\index{Reputation Tiers!Veteran}
\item[Tier IV -- Paragon (151--220 XP)] Movers and shakers; rivals emerge.\index{Reputation Tiers!Paragon}
\item[Tier V -- Mythic (221+ XP)] Legendary status; kingdoms respond.\index{Reputation Tiers!Mythic}
\end{description}

\subsection{Player Archetypes}
\label{subsec:archetypes-ref}
\index{Appendices!Player Archetypes}

\begin{description}
\item[Solo] 70--90\% self investment; minimal followers/assets.\index{Character Build!Solo}
\item[Mixed] 50--65\% self; balanced with followers/assets.\index{Character Build!Mixed Player}
\item[Mastermind] 25--40\% self; focuses on networks and followers.\index{Character Build!Mastermind}
\end{description}

\section{Gamemaster Guidance}
\label{sec:gm-reference}
\index{Appendices!GM Guidance}

\subsection{Session Preparation Checklist}
\label{subsec:session-prep-ref}
\index{Appendices!Session Preparation}

\begin{itemize}
\item Review previous session notes and unresolved threads.
\item Set initial SB budget (4 + character tiers).
\item Prepare key scenes and opposition.
\item Have consequence ideas ready for common actions.
\item Check ongoing clocks and faction status.
\item Prepare travel routes if journey expected.
\end{itemize}

\subsection{Adjudication Principles}
\label{subsec:adjudication-ref}
\index{Appendices!Adjudication}

\begin{itemize}
\item \textbf{Fiction First}: Mechanics serve the narrative, not replace it.
\item \textbf{Fail Forward}: Even failures should advance the story.
\item \textbf{Player Agency}: Offer choices rather than impose outcomes.
\item \textbf{Transparent Costs}: Clearly communicate risks and stakes.
\item \textbf{Collaborative Spirit}: Work with players to create compelling fiction.
\end{itemize}

\subsection{Pacing Tools}
\label{subsec:pacing-tools-ref}
\index{Appendices!Pacing}

\begin{itemize}
\item Use clocks to create urgency and track progress.
\item Vary scene intensity between high and low stakes.
\item Include downtime for character development.
\item Balance action, investigation, and social scenes.
\item Use travel sequences for world-building and random encounters.
\end{itemize}

\section{Common Rules Questions}
\label{sec:rules-questions}
\index{Appendices!Rules Questions}

\subsection{Core Mechanic Clarifications}
\label{subsec:core-clarifications}
\index{Appendices!Core Clarifications}

\textbf{Q: Can players re-roll 1s to remove Story Beats?}\\
A: No. Re-rolling 1s does not remove SB already generated. If re-rolled dice show 1 again, they generate additional SB.\index{Story Beats!Re-rolling}

\textbf{Q: When does a miss or partial success award Boons?}\\
A: Only when all three conditions are met: procedure followed, stakes stated, and consequence lands immediately.\index{Boons!Award Conditions}

\textbf{Q: Can players assist each other on every action?}\\
A: Yes, but total assist dice are capped at +3 from all sources combined.\index{Mechanics!Assist Maximum}

\subsection{Magic System Questions}
\label{subsec:magic-questions}
\index{Appendices!Magic Questions}

\textbf{Q: Can a character use multiple magic paths?}\\
A: Yes, but each path has its own tracking (Backlash, Obligation, Symbol states). Specializing is more efficient.\index{Magic!Mixing Paths}

\textbf{Q: How does Crack the Seal work for Invokers?}\\
A: Convert a ritual to instant casting by setting the Symbol to \emph{Compromised} and marking +2/+3 Obligation.\index{Invokers!Crack the Seal}

\textbf{Q: What happens when a Patron's Obligation clock fills?}\\
A: The GM resolves the debt in-fiction through service demands, omens, or narrative consequences.\index{Magic!Obligation Resolution}

\subsection{Combat and Conflict Questions}
\label{subsec:combat-questions}
\index{Appendices!Combat Questions}

\textbf{Q: How does the Over-Stack rule work?}\\
A: If the party enters a scene with $\geq$ 3 structural advantages, either start one challenge at +1 difficulty \emph{or} bank +1 SB.\index{Combat!Over-Stack}

\textbf{Q: Can players spend Boons to improve Position?}\\
A: Yes. 1 Boon improves Position by 1 step for the current action.\index{Combat!Position Improvement}

\textbf{Q: How does harm recovery work?}\\
A: Minor clears with rest; moderate requires medical treatment (DV 2); severe needs extended care (DV 3); critical requires major intervention.\index{Combat!Harm Recovery}

\section{Regional Quick Reference}
\label{sec:regional-reference}
\index{Appendices!Regional Reference}

\subsection{Major Regions and Themes}
\label{subsec:regions-themes}
\index{Appendices!Regions}

\begin{description}
\item[Acasia] Broken marches, curses, lawless territory.\index{Acasia}
\item[Aeler] Underground vaults, dwarven culture, engineering.\index{Aeler}
\item[Ecktoria] Imperial remnants, bureaucracy, coinhouses.\index{Ecktoria}
\item[Kahfagia] Maritime trade, lantern-law, convoys.\index{Kahfagia}
\item[Mistlands] Bells, wards, supernatural boundaries.\index{Mistlands}
\item[Silkstrand] Trade hub, intrigue, Acasia's only major city.\index{Silkstrand}
\item[Vhasia] Fractured sun, political fragmentation.\index{Vhasia}
\item[Viterra] Last kingdom, river-based power.\index{Viterra}
\item[Valewood] Forest empire, natural magic.\index{Valewood}
\item[Ykrul] Steppe nomads, wolf standards.\index{Ykrul}
\item[Zakov] Salt and serpent, criminal syndicates.\index{Zakov}
\end{description}

\subsection{Key Geographical Features}
\label{subsec:geography-ref}
\index{Appendices!Geography}

\begin{itemize}
\item \textbf{Amaranthine Sea}: Western sea, major trade routes.\index{Geography!Amaranthine Sea}
\item \textbf{Dolmis Sea}: Inner sea, island networks.\index{Geography!Dolmis Sea}
\item \textbf{Astroegro Straits}: Crucial maritime chokepoint.\index{Geography!Astroegro Straits}
\item \textbf{Belworth River}: Major river system, boundary between regions.\index{Geography!Belworth River}
\item \textbf{Aelerian Mountains}: Extensive underground networks.\index{Geography!Aelerian Mountains}
\end{itemize}

\section{Campaign Management Tools}
\label{sec:campaign-tools}
\index{Appendices!Campaign Tools}

\subsection{Session Log Template}
\label{subsec:session-log-template}
\index{Appendices!Session Log}

\begin{center}
\feTableStart
\begin{tabularx}{\linewidth}{@{}p{4cm} Y @{}}
\toprule
\textbf{Session Element} & \textbf{Notes} \\
\midrule
Session Date & \\
Players Present & \\
Major Objectives & \\
Key Scenes & \\
Story Beats Generated & \\
Boons Awarded/Spent & \\
Clocks Advanced/Completed & \\
XP Awards & \\
Downtime Activities & \\
Next Session Hooks & \\
\bottomrule
\end{tabularx}
\feTableEnd
\end{center}

\subsection{Campaign Clock Examples}
\label{subsec:campaign-clocks}
\index{Appendices!Campaign Clocks}

\begin{description}
\item[\textbf{Faction Rivalry} (8 segments)] Tracks escalating conflict between major powers.\index{Clocks!Faction Rivalry}
\item[\textbf{Ancient Curse} (6 segments)] Progress of a regional supernatural affliction.\index{Clocks!Ancient Curse}
\item[\textbf{Imperial Collapse} (10 segments)] Decline of a major governing power.\index{Clocks!Imperial Collapse}
\item[\textbf{Magical Cataclysm} (8 segments)] Buildup to a reality-altering event.\index{Clocks!Magical Cataclysm}
\item[\textbf{Trade War} (6 segments)] Economic conflict affecting multiple regions.\index{Clocks!Trade War}
\end{description}

\subsection{Adventure Structure Template}
\label{subsec:adventure-template}
\index{Appendices!Adventure Template}

\textbf{Standard Three-Act Structure:}
\begin{enumerate}
\item \textbf{Introduction}: Establish situation; introduce key NPCs and locations.
\item \textbf{Development}: 2--3 challenges that advance the main objective.
\item \textbf{Climax}: Major confrontation or resolution point.
\item \textbf{Resolution}: Consequences and setup for future adventures.
\end{enumerate}

\textbf{Alternative Structures:}
\begin{itemize}
\item \textbf{Hex Crawl}: Exploration-focused with multiple points of interest.
\item \textbf{Mystery}: Investigation-driven with clue accumulation.
\item \textbf{Siege}: Defense-focused with resource management.
\item \textbf{Journey}: Travel-based with episodic encounters.
\end{itemize}

\section{Troubleshooting Common Issues}
\label{sec:troubleshooting-ref}
\index{Appendices!Troubleshooting}

\subsection{Player Engagement Issues}
\label{subsec:engagement-issues}
\index{Appendices!Engagement Issues}

\textbf{Issue: Players are passive or hesitant}
\begin{itemize}
\item \textbf{Solutions}: Use leading questions, offer clear options, create immediate stakes.
\item Provide obvious hooks and direct incentives for action.
\item Use NPCs to demonstrate active approaches.
\item Reward proactive play with narrative advantages.
\end{itemize}

\textbf{Issue: Rules discussions slow the game}
\begin{itemize}
\item \textbf{Solutions}: Make quick rulings; note for later review; keep momentum.
\item Establish “ruling now, researching later” policy.
\item Designate one player as rules reference to minimize lookups.
\item Use standardized procedures for common actions.
\end{itemize}

\subsection{Balance and Challenge Issues}
\label{subsec:balance-issues}
\index{Appendices!Balance Issues}

\textbf{Issue: Encounters are too easy or too hard}
\begin{itemize}
\item \textbf{Solutions}: Adjust opposition on the fly; use SB to modulate difficulty.
\item Remember that Position and Effect can be adjusted situationally.
\item Use environmental factors to change challenge levels.
\item Allow creative solutions to bypass straight combat.
\end{itemize}

\textbf{Issue: Magic feels overpowered or underpowered}
\begin{itemize}
\item \textbf{Solutions}: Ensure proper Backlash and Obligation application.
\item Remember that high-DV spells carry significant risks.
\item Use countermagic and magical opposition when appropriate.
\item Ensure non-magical characters have meaningful contributions.
\end{itemize}

\subsection{Tracking and Administration Issues}
\label{subsec:tracking-issues}
\index{Appendices!Tracking Issues}

\textbf{Issue: Too much bookkeeping slows play}
\begin{itemize}
\item \textbf{Solutions}: Simplify tracking to essential elements; delegate to players.
\item Use abstract ranges and conditions rather than precise measurements.
\item Focus on narrative consequences rather than numerical modifiers.
\item Use index cards or digital tools for complex tracking.
\end{itemize}

\textbf{Issue: Players forget abilities or resources}
\begin{itemize}
\item \textbf{Solutions}: Provide quick reference sheets; use visual aids.
\item Create character-specific reminder cards.
\item Use recap sessions to review capabilities.
\item Encourage players to maintain updated character sheets.
\end{itemize}

\section{Advanced Play Techniques}
\label{sec:advanced-techniques}
\index{Appendices!Advanced Techniques}

\subsection{Narrative-First Adjudication}
\label{subsec:narrative-adjudication}
\index{Appendices!Narrative Adjudication}

\begin{itemize}
\item Ask “What happens next?” rather than “What’s the rule?”
\item \textbf{Use the Fiction}: Let the narrative dictate mechanical outcomes.
\item \textbf{Embrace Improvisation}: Create rulings that serve the story.
\item \textbf{Collaborative World-Building}: Involve players in creating details.
\item \textbf{Consequence-Driven Play}: Ensure every action has meaningful results.
\end{itemize}

\subsection{Pacing and Rhythm Management}
\label{subsec:pacing-management}
\index{Appendices!Pacing Management}

\textbf{Scene Pacing:}
\begin{itemize}
\item Vary intensity between high-action and quiet moments.
\item Use clocks to create natural endpoints.
\item Alternate between player-driven and GM-driven scenes.
\item Include breather moments for character development.
\end{itemize}

\textbf{Campaign Rhythm:}
\begin{itemize}
\item Balance episodic adventures with ongoing arcs.
\item Use downtime effectively between major events.
\item Vary the scope of challenges (personal, local, regional, global).
\item Include both planned and emergent story elements.
\end{itemize}

\subsection{Player Spotlight Management}
\label{subsec:spotlight-management}
\index{Appendices!Spotlight Management}

\begin{itemize}
\item \textbf{Rotate Focus}: Ensure each character gets meaningful scenes.
\item \textbf{Personal Arcs}: Develop individual character stories.
\item \textbf{Group Dynamics}: Create situations that require teamwork.
\item \textbf{Specialization Respect}: Allow experts to shine in their domains.
\item \textbf{Shared Moments}: Include scenes that develop group bonds.
\end{itemize}

\clearpage
% =========================
% Fate's Edge — Tags Reference
% =========================
% Include in main with: \input{tags-reference.tex}
% Uses only standard longtable/booktabs; no custom macros required.

\section*{Spell [TAGS] Reference}
\label{sec:spell-tags}

\setlength{\LTpre}{6pt}
\setlength{\LTpost}{6pt}
\renewcommand{\arraystretch}{1.08}

\small
\begin{longtable}{@{} l l p{4.2cm} c p{6.5cm} p{2.2cm} p{3.2cm} p{2.2cm} @{}}%
\toprule
\textbf{[TAG]} & \textbf{Category} & \textbf{Gated By} & \textbf{DV} & \textbf{Effect (summary)} & \textbf{Duration} & \textbf{Cost/Risk} & \textbf{Counters} \\
\midrule
\endfirsthead
\toprule
\textbf{[TAG]} & \textbf{Category} & \textbf{Gated By} & \textbf{DV} & \textbf{Effect (summary)} & \textbf{Duration} & \textbf{Cost/Risk} & \textbf{Counters} \\
\midrule
\endhead
\midrule
\multicolumn{8}{r}{\emph{Continued on next page}}\\
\endfoot
\bottomrule
\endlastfoot

% --- CONTROL & RESTRAINT ---
\textbf{BIND}      & Control/Restraint & Ritual: Gate; Skill: Command+Arcana & 3--4 & Fix target in place or to locus; may start/advance a \emph{Leash}. & 1 beat (sustain) & SB on 1s; sustain = 1 Fatigue/beat & DISPEL, SEVER \\
\textbf{SEAL}      & Control/Boundary  & Ritual: Wards; Talent: Sealwright   & 3--5 & Close portal/container; lock threshold or suppress effect. & Scene/clock & Obligation (Ritual) or SB & UNSEAL, UNWARD \\
\textbf{UNSEAL}    & Control/Boundary  & Skill: Arcana+Wits                   & 3--4 & Open sealed thing without damage (if keyed). & Instant & SB; alarms possible & SEAL, WARD \\
\textbf{COMMAND}   & Control/Will      & Ritual: Dominion; Skill: Presence    & 3--5 & Issue short imperative to lesser mind/spirit. & 1 beat & SB; +1 DV vs strong wills & RESIST, COUNTER \\
\textbf{COMPEL}    & Control/Geas      & Skill: Geasa; Ritual: Oath           & 4--5 & Bind stated taboo/trigger (attach 4-seg clock). & Arc/clock & Obligation or SB on failures & UNGEAS, DISPEL \\
\textbf{LEASH}     & Summoning Control & Ritual: Gate; Talent: Binder         & 2--4 & Establish/advance control track for a summoned entity. & While active & SB on strain triggers & SEVER, RIVAL \\
\textbf{SEVER}     & Control/Break     & Skill: Arcana or Fate                & 4--5 & Cut tethers (Leashes, bindings, glamours). & Instant & Backlash by Element & REWEAVE \\
\textbf{WARD}      & Protection/Zone   & Ritual: Wards; Skill: Runecraft      & 2--4 & Block/filter entities/effects by tag. Outsiders: DV = Cap to cross. & Scene / 3 beats & Obligation or SB & UNWARD, DISPEL \\
\textbf{UNWARD}    & Protection/Break  & Skill: Arcana+Tools                  & 3--4 & Suppress/unpick a ward’s key. & 1 beat & SB; noise/trace & WARD \\
\textbf{REFLECT}   & Protection/Meta   & Talent: Counter-Circle               & 4--5 & Turn next targeted effect back on source. & 1 use / 1 beat & SB; risky on Partial & COUNTER, DISPEL \\

% --- META-MAGIC ---
\textbf{DISPEL}    & Meta/End Effect   & Skill: Arcana; Ritual: Unravel       & 2--5 & End ongoing magical effect (DV by scale). & Instant & SB; collateral & — \\
\textbf{COUNTER}   & Meta/Interrupt    & Skill: Arcana (Reaction)             & 3--4 & Interrupt casting/ritual in its window. & Instant & SB; on Miss GM +1 SB & — \\
\textbf{AMPLIFY}   & Meta/Boost        & Talent: Focused Weave                & +1 DV & Increase Effect one step on current working. & Instant & +1 SB or +1 Obligation & COUNTER \\
\textbf{MITIGATE}  & Meta/SB Control   & Talent/Asset: “Mitigate SB”          & —     & Convert up to 2 SB from this cast to a strain clock. & Instant & 1 Fatigue or asset tick & — \\
\textbf{EXTEND}    & Meta/Duration     & Talent: Sustainer                     & +1 DV & Double default duration. & As set & +1 Fatigue/beat & DISPEL \\
\textbf{REWEAVE}   & Meta/Repair       & Skill: Arcana                         & 3--4 & Patch/stabilize a faltering effect/clock. & Instant & SB or component & DISPEL \\

% --- MOVEMENT & SPACE ---
\textbf{LEVITATE}  & Movement/Air      & Element: Air; Art: Kinesis           & 3--4 & Lift/hold target or self at Near height. & Sustain & SB: drift/drop & GROUNDED, COUNTER \\
\textbf{PULL}      & Movement/Vector   & Element: Air/Earth                   & 2--3 & Draw object/creature toward locus. & Instant & SB: collide/entangle & ANCHOR \\
\textbf{PUSH}      & Movement/Vector   & Element: Air/Force                   & 2--3 & Shove/knockback one band. & Instant & SB: spin/overpush & ANCHOR \\
\textbf{ANCHOR}    & Space/Lock        & Element: Earth/Fate                  & 3--4 & Fix object/zone; resists movement/banish. & Scene & SB: stuck at wrong time & SEVER \\
\textbf{TRANSPORT} & Space/Move Many   & Ritual: Passage; Art: Wayfinding     & 4--5+ & Move targets across bands/locations. & Instant & Prep/Obligation; SB: scatter & WARD, COUNTER \\
\textbf{TELEPORT}  & Space/Blink       & Element: Fate/Air                    & 5+    & Instant relocate (LoS or keyed sigil). & Instant & High SB on Partial & WARD, COUNTER \\

% --- PERCEPTION & INFORMATION ---
\textbf{SENSE}     & Perception        & Skill: Insight/Arcana                & 2--3 & Detect presence of a named tag/element. & Instant & SB: false positives & SHROUD \\
\textbf{REVEAL}    & Perception        & Element: Light/Fate                  & 3--4 & Unveil hidden/glamoured things. & 1 beat & SB: overexposure & SHROUD, GLAMOUR \\
\textbf{SCRY}      & Divination        & Ritual: Sight; Skill: Lore           & 3--5 & Observe distant place/target via focus. & Scene / 3 beats & SB: noticed/scry-back & VEIL, COUNTER \\
\textbf{DIVINE}    & Divination/Omens  & Element: Fate                        & 3--4 & Ask likely outcome/path. & Instant & SB: omen debt/paradox & — \\
\textbf{MARK}      & Tracking          & Talent: Hunter’s Sigil               & 2--3 & Tag target for later SENSE/SCRY. & Arc / 3 scenes & SB on application & CLEANSE \\

% --- PROTECTION & BUFFS ---
\textbf{PROTECT}   & Protection/Shield & Element: Water/Earth                 & 2--4 & Reduce/deflect next harm or tag. & 1–3 beats / 1 use & SB: thematic backlash & DISPEL, PIERCE \\
\textbf{FORTIFY}   & Protection/Buff   & Element: Earth/Life                  & 3--4 & Boost resilience; improve Position/Effect. & 3 beats & SB: rigidity/speed loss & DISPEL \\
\textbf{ABSORB}    & Protection/Soak   & Talent: Spellguard                   & 4--5 & Convert incoming harm to SB or Fatigue. & Instant & Fatigue + SB & PIERCE \\
\textbf{HASTE}     & Time/Boost        & Element: Air/Fate                    & 3--4 & Speed one ally; extra Move or +Effect. & 1–3 beats & SB: tunnel vision & DISPEL \\
\textbf{SLOW}      & Time/Hinder       & Element: Water/Fate                  & 3--4 & Impose drag; worsen Position or remove Move. & 1–3 beats & SB: spillover & DISPEL, RESIST \\

% --- WEAKEN & DEBILITATE ---
\textbf{WEAKEN}    & Debuff            & Element: Fate/Death                  & 3--4 & Lower Attribute/Skill effectiveness one step. & 1–3 beats & SB: misfire to ally & DISPEL, RESIST \\
\textbf{ENTANGLE}  & Debuff/Control    & Element: Earth/Plant                 & 2--3 & Restrain; worsen Position or steal action. & 1–2 beats & SB: self-snag & SEVER, BURN \\
\textbf{SILENCE}   & Debuff/Nullify    & Element: Air/Shadow                  & 3--4 & Suppress sound/casting words in zone/target. & 1–3 beats & SB: dull senses & BREAK, COUNTER \\
\textbf{BLIND}     & Debuff/Sense      & Element: Shadow/Light                & 3--4 & Obscure sight of target/zone. & 1–2 beats & SB: friendly fire & REVEAL, CLEANSE \\

% --- ELEMENTAL MANIPULATION ---
\textbf{IGNITE}    & Fire/Heat         & Element: Fire                        & 2--3 & Set alight; create burning hazard. & Instant / 1–3 beats & SB: spread/scorch & DOUSE \\
\textbf{DOUSE}     & Water/Quench      & Element: Water                       & 2--3 & Extinguish, cool, suppress heat. & Instant & SB: steam/slick & IGNITE \\
\textbf{STONE}     & Earth/Shape       & Element: Earth                       & 3--4 & Raise wall, seal crack, shape masonry. & Scene/clock & SB: collapse/rigidity & SEVER, SHATTER \\
\textbf{SHOCK}     & Air/Lightning     & Element: Air                         & 2--4 & Stun/overload; arc between targets. & Instant & SB: backfeed & GROUND \\
\textbf{FREEZE}    & Water/Cold        & Element: Water                       & 3--4 & Ice over; slow or trap. & 1–3 beats & SB: brittle snap & MELT \\
\textbf{MELT}      & Fire/Heat         & Element: Fire                        & 2--3 & Thaw/liquefy; clear FREEZE. & Instant & SB: runoff & FREEZE \\
\textbf{WIND}      & Air/Control       & Element: Air                         & 2--3 & Gusts, deflect missiles, carry sound. & 1–3 beats & SB: scatter/noise & ANCHOR \\
\textbf{FLOOD}     & Water/Volume      & Element: Water                       & 4--5 & Surge of water; zone hazard/knockdown. & 1–3 beats & SB: collateral & DRAIN, FREEZE \\
\textbf{DRAIN}     & Water/Void        & Element: Water/Earth                 & 3--4 & Clear liquids/pressure from zone. & Instant & SB: sinkhole & — \\
\textbf{LIGHT}     & Radiance          & Element: Fire/Light                  & 2--3 & Illuminate, reveal, or blind (high Effect). & 1–3 beats & SB: beacon/attention & SHROUD \\
\textbf{SHADOW}    & Obscuration       & Element: Shadow/Death                & 2--3 & Deepen darkness; hide edges/tracks. & 1–3 beats & SB: things watch & LIGHT, REVEAL \\

% --- LIFE, DEATH & SPIRIT ---
\textbf{HEAL}      & Life/Repair       & Element: Life                        & 3--4 & Mend living harm one step; stabilize. & Instant & SB: fever/overgrowth & WITHER \\
\textbf{RESTORE}   & Structure/Mend    & Skill: Craft; Ritual: Repair         & 3--4 & Remove \texttt{COMPROMISED} from gear/constructs. & Scene / 1 beat & Components or SB & SHATTER \\
\textbf{CLEANSE}   & Purify            & Element: Water/Life                  & 2--3 & Remove poisons/blights/marks. & Instant & SB: brief weakness & CURSE \\
\textbf{CURSE}     & Bane              & Ritual: Malediction; Skill: Hex      & 3--5 & Attach hostile tag/clock to target. & Arc/clock & Obligation or SB & CLEANSE, DISPEL \\
\textbf{BLESS}     & Boon              & Ritual: Benediction                  & 2--3 & Grant favorable tag (luck, favor, ward-key). & 3 beats / scene & Minor Obligation & DISPEL \\
\textbf{BANISH}    & Spirit/Expel      & Ritual: Gate; Skill: Exorcism        & 3--5 & Drive outsider/spirit across threshold (DV = Cap). & Instant & SB: rage/echo & ANCHOR, WARD \\
\textbf{SUMMON}    & Spirit/Call       & Ritual: Gate                         & 3--5 & Call named class of entity; start \emph{Leash}. & Scene/contract & Obligation; SB: hungry guest & WARD, COUNTER \\
\textbf{DREAM}     & Mind/Oneiric      & Element: Death/Dreams                & 3--4 & Enter/influence sleeping mind/realm. & Scene & SB: nightmares/bleed & WAKE, WARD \\
\textbf{WAKE}      & Mind/Counter      & Skill: Medicine or Spirit            & 2--3 & Break sleep/charm; end DREAM. & Instant & SB: shock/disorientation & DREAM \\
\textbf{NIGHTMARE} & Mind/Fear         & Element: Death/Dreams                & 3--4 & Inflict fear/visions; fill small clock. & 1–3 beats & SB: echo lingers & CLEANSE, STEEL \\

% --- FATE & FORTUNE ---
\textbf{FATELOCK}  & Fate/Anchor       & Element: Fate; Talent: Oaths         & 4--5 & Fix outcome boundary (“this door will hold”). & Scene/clock & Paradox SB later & UNWEAVE, LUCK \\
\textbf{LUCK}      & Fortune/Shift     & Element: Luck                        & 3--4 & Tilt odds; convert 1 Miss to 1 Success (once). & 1 beat & SB: side-effect elsewhere & FATE \\
\textbf{OMEN}      & Fortune/Flag      & Element: Fate/Luck                   & 2--3 & Mark looming twist the GM may cash later. & Arc & GM banks +1 SB & — \\
\textbf{PROPHECY}  & Fate/Foretell     & Ritual: Oracle                       & 4--5 & Declare true but costly future clause. & Campaign & SB bank + omen costs & FATELOCK \\

% --- ILLUSION & DECEPTION ---
\textbf{GLAMOUR}   & Illusion/Seeming  & Element: Air/Shadow                  & 3--4 & Alter appearance/sound/scent; not substance. & 1–3 beats & SB: seams show & REVEAL, DISBELIEVE \\
\textbf{VEIL}      & Illusion/Hide     & Element: Shadow/Water                & 2--3 & Hide subject/area from casual notice. & 1–3 beats & SB: blindspot & REVEAL, SENSE \\
\textbf{PHANTOM}   & Illusion/Construct& Talent: Image-Craft                  & 3--4 & Create interactive illusion (no weight). & 1–3 beats & SB: bleed/noise & REVEAL, DISBELIEVE \\

% --- SOUND, WORD, LAW ---
\textbf{NAME}      & Word/True Name    & Talent: Name-Caller                  & 4--5 & Speak true name to compel or reveal. & Instant / 1 beat & SB: name echoes & SILENCE, COUNTER \\
\textbf{OATH}      & Word/Binding      & Skill: Geasa; Ritual: Pact           & 3--4 & Bind parties to terms; create oath clock. & Arc & Shared Obligation & UNGEAS, FATELOCK \\
\textbf{UNGEAS}    & Word/Release      & Skill: Geasa+Utaran Law (or Ritual)  & 4--5 & Lawful release or loophole clause. & Instant & SB: fallout & OATH, GEAS \\

% --- TERRAIN & STRUCTURE ---
\textbf{BRIDGE}    & Terrain/Create    & Element: Earth/Water/Air             & 3--4 & Create traversable surface/path. & Scene & SB: sway/crack & SHATTER, FLOOD \\
\textbf{SHATTER}   & Terrain/Break     & Element: Sound/Stone                 & 3--4 & Rupture brittle/strained matter. & Instant & SB: debris & STONE \\
\textbf{SANCTIFY}  & Place/Threshold   & Ritual: Consecration                 & 3--4 & Make area inhospitable to a named tag. & Scene/arc & Obligation & PROFANE, UNWARD \\
\textbf{PROFANE}   & Place/Defile      & Ritual: Desecration                  & 3--4 & Flip sanctity; empower rival forces. & Scene/arc & Obligation; social fallout & SANCTIFY \\

% --- EDGE CASES & UTILITY ---
\textbf{TETHER}    & Utility/Link      & Element: Fate                        & 2--3 & Link two objects for later PULL/SENSE. & Arc & SB on stretch & SEVER \\
\textbf{STORE}     & Utility/Charge    & Talent: Capacitor                    & 3--4 & Bank 1–2 successes in a vessel (once). & Scene/arc & SB if overfilled & DISPEL \\
\textbf{TRANSFER}  & Utility/Hand-off  & Skill: Arcana+Tools                  & 3--4 & Move an ongoing effect between hosts. & Instant & SB: spill & DISPEL \\

\end{longtable}
\normalsize

\paragraph{Usage Notes}
\begin{itemize}
  \item \textbf{DV guidance:} 2 small/local, 3 scene-scale, 4 big swing, 5+ set-piece/ritual.
  \item \textbf{Duration defaults:} Buffs $\approx$ 3 beats; areas 1 beat. Sustaining costs 1 Fatigue/beat.
  \item \textbf{Costs:} “SB” = Story Beats generated/at risk; \emph{Obligation} applies to \textbf{Ritual} sources; some tags list sustain/asset ticks.
  \item \textbf{Counters:} A listed counter doesn’t auto-succeed—resolve with the appropriate action (often \texttt{COUNTER}, \texttt{DISPEL}, or fiction).
  \item \textbf{Gating:} “Gated By” lists \emph{typical} prerequisites—swap in specific Patrons, Arts, or Talents to taste.
\end{itemize}

% =========================
% QUICK REFERENCE
% =========================
\section{Quick Reference: Fatigue \& Harm}
\label{sec:quick-ref-fatigue-harm}

\subsection*{Fatigue Track}
\begin{tabular}{@{}ll@{}}
\textbf{Boxes} & Body \\
\textbf{Clear on} & Short Rest (-2), Full Night (all), or [HEALED] (all) \\
\end{tabular}

\subsection*{Overflow}
\begin{tabular}{@{}lll@{}}
\textbf{Trigger} & \textbf{Result} & \textbf{Then} \\
Fatigue track fills & Harm +1 level & Clear all Fatigue \\
\end{tabular}

\subsection*{Obligation $\rightarrow$ Fatigue}
\begin{tabular}{@{}lll@{}}
\textbf{Severity} & \textbf{Fatigue} & \textbf{Note} \\
Low & +1 & Nagging strain \\
Standard & +2 & Significant cost \\
High & +3 & Brutal, only for dire rites \\
\end{tabular}

\chapter{Weapons \& Armor Quick Reference}
\label{app:weapons-armor}
\index{Weapons}\index{Armor}

\section{Weapons by Weight Class}
\label{app:weapons-weight}

\subsection*{Costs \& Baselines}
\begin{itemize}
  \item \textbf{Light (4 XP)} — quick, concealable.
  \item \textbf{Medium (8 XP)} — balanced, battlefield mainstays.
  \item \textbf{Heavy (12 XP)} — punishing, slow to bring to bear.
\end{itemize}

\subsection*{Melee Modifiers (by Range Band)}
\begin{center}
\begin{tabular}{llll}
\toprule
\textbf{Weight} & \textbf{Close} & \textbf{Near} & \textbf{Notes} \\
\midrule
Light & +2 dice & +1 die & Fast in tight quarters \\
Medium & +1 die & +2 dice & \emph{Set} once/scene or -1 die on first attack \\
Heavy & -1 die  & +3 dice & \emph{Set} once/scene or -2 dice on first attack \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Ranged Modifiers \& Tempo}
\label{app:ranged-tempo}
\begin{center}
\begin{tabular}{lllll}
\toprule
\textbf{Weight} & \textbf{Tempo} & \textbf{Close} & \textbf{Near} & \textbf{Far} \\
\midrule
Light (4 XP) & \textbf{Fast} & Risky (Engaged) & +1 die & — \\
Medium (8 XP) & \textbf{Standard} & Desperate (Engaged) & +2 dice & +1 die \\
Heavy (12 XP) & \textbf{Slow} & Desperate (Engaged) & +1 die & +3 dice \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Tempo shorthand (see \S\ref{sec:ranged}):}
\begin{itemize}
  \item \textbf{Fast} — Move \emph{and} Shoot; reload is incidental. Risky at Close while Engaged.
  \item \textbf{Standard} — \emph{Move or Shoot}; partial-turn reload. May \emph{Aim} for +1 die or +1 Effect.
  \item \textbf{Slow} — Requires Set/Brace; full-turn reload; cannot Move and Shoot same round.
\end{itemize}

\subsection*{Common Examples (Non-exhaustive)}
\begin{itemize}
  \item \textbf{Light Melee:} dagger, hatchet, cudgel, short-spear, main-gauche.
  \item \textbf{Medium Melee:} arming sword, war pick, mace, battle-axe, spear.
  \item \textbf{Heavy Melee:} greatsword, poleaxe, maul, glaive, long-spear.
  \item \textbf{Light Ranged (Fast):} throwing knives, javelins, sling, shortbow.
  \item \textbf{Medium Ranged (Standard):} warbow, hand crossbow, light crossbow.
  \item \textbf{Heavy Ranged (Slow):} greatbow, arbalest, pavise-rest crossbow.
\end{itemize}

\subsection*{Weapon Tags (Optional, +4 XP each, max 2)}
\label{app:weapon-tags}
\index{Weapons!Tags}
\begin{description}
  \item[\textbf{Reach}] Ignore Engage cost once/round; foes suffer +1 DV at Close.
  \item[\textbf{Close}] No penalty at grappling distance; safer to draw in melee.
  \item[\textbf{Accurate}] +1 die when you \emph{Aim}.
  \item[\textbf{Brutal}] +1 Effect on a 6+, but the roll generates +1 SB if any \textbf{1} appears.
  \item[\textbf{Hook}] On hit, may shift target to worse Position instead of extra Effect.
  \item[\textbf{Concealable}] No penalty to smuggle; draw is incidental.
  \item[\textbf{Quickdraw}] First attack you make this scene is +1 die.
  \item[\textbf{Two-Handed}] Stronger leverage; cannot benefit from shields.
  \item[\textbf{Off-Hand}] When dual-wielding, +1 die to one \emph{Attack} per round.
\end{description}

\section{Shields (Optional)}
\label{app:shields}
\index{Shields}
\begin{center}
\begin{tabular}{llll}
\toprule
\textbf{Shield} & \textbf{XP} & \textbf{Benefit (per round)} & \textbf{Tradeoff} \\
\midrule
Buckler & 4 & +1 die to \emph{Defend} vs melee \emph{or} +1 DV once & Off-hand occupied \\
Heater  & 8 & +1 die to \emph{Defend} and convert 1 Harm$\rightarrow$1 Fatigue once & -1 die to your Ranged attacks \\
Pavise  & 12 & \emph{Plant} (action): heavy cover cone; allies may use & Slow to move; counts as bulky \\
\bottomrule
\end{tabular}
\end{center}

\section{Armor}
\label{app:armor}
\index{Armor!Conversion}
Armor converts incoming Harm into Fatigue (see \S\ref{sec:health-fatigue-harm-rev}). Choose one suit.

\begin{center}
\begin{tabular}{llll}
\toprule
\textbf{Armor} & \textbf{XP} & \textbf{Conversion} & \textbf{Penalties} \\
\midrule
Light  & 4  & Up to 1 Harm $\rightarrow$ 1 Fatigue & — \\
Medium & 8  & Up to 2 Harm $\rightarrow$ 1 Fatigue & -1 die to physical skills \\
Heavy  & 12 & Up to 3 Harm $\rightarrow$ 2 Fatigue & -2 dice to physical skills; no sprint in rough \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Notes:}
\begin{itemize}
  \item Apply conversion before marking Fatigue. If this fills the Fatigue track, resolve Fatigue$\rightarrow$Harm conversion and clear Fatigue (see \S\ref{sec:health-fatigue-harm-rev}).
  \item Conversion is per incoming instance of Harm; you may still \emph{Resist} to reduce level first.
\end{itemize}

\section{Condition \& Upkeep}
\label{app:condition}
\index{Weapons!Condition}\index{Armor!Condition}
\begin{description}
  \item[\textbf{Neglected}] Weapons: -1 die until maintained. Armor: conversion reduced by 1 step (Light=0; Medium treats 2$\rightarrow$1F; Heavy treats 3$\rightarrow$2F).
  \item[\textbf{Compromised}] Weapons: -1 die on your first attack each exchange. Armor: no conversion.
\end{description}
\noindent\emph{Maintenance:} A Short Rest with tools removes \textbf{Neglected}. A scene of work or a Smith removes \textbf{Compromised}.

\section{At-a-Glance Options (Ranged)}
\label{app:ranged-at-a-glance}
\index{Combat!Ranged!Options}
\begin{itemize}
  \item \textbf{Aim:} +1 die \emph{or} +1 Effect next shot (lost if you Move or are disrupted).
  \item \textbf{Volley:} Spend extra ammo for +1 die (max +2).
  \item \textbf{Suppress:} Zone under fire; foes there act at -1 die or Limited Effect until they shift cover.
  \item \textbf{Overwatch:} Ready a Risky shot on a clear trigger crossing your line.
\end{itemize}

% Index shortcuts
\index{Weapons!Light}\index{Weapons!Medium}\index{Weapons!Heavy}
\index{Ranged!Fast}\index{Ranged!Standard}\index{Ranged!Slow}
\index{Armor!Light}\index{Armor!Medium}\index{Armor!Heavy}
\index{Fatigue!Conversion}\index{Harm!Conversion}

\clearpage

\chapter{The Whispered Grimoire (Rituals)}

\section*{The Ritual of Exorcism}

When possession takes root, when the shadow has grown too deep, there remains one path to salvation: the Ritual of Exorcism. But know this—exorcism is not a simple spell cast and forgotten. It is warfare. It is negotiation. It is sometimes surrender.

\textbf{Purpose:} To forcibly remove a possessing entity from a possessed subject

\textbf{Primary Caster:} Spirit 3+, Lore 3+, Insight 2+

\textbf{Helpers:} 1+ participants with relevant skills (Lore, Spirit, Insight, Command, Presence)

\textbf{The Exorcism Encounter:}

Exorcism functions as a special encounter with its own contested clock mechanics. The ritual creates a contested space where the caster's will battles the possessing entity for control of the subject's soul.

\textbf{Exorcism Clock:} 2 + Subject's Tier in segments (minimum 4 segments, maximum 8 segments)

Each segment represents a phase of spiritual combat requiring:
\begin{enumerate}
\item \textbf{Invocation Phase (Spirit + Lore):} Primary caster and helpers invoke sacred power and knowledge
\item \textbf{Confrontation Phase (Presence + Insight or Command):} Direct mental/spiritual assault on the possessing entity
\item \item \textbf{Binding Phase (Wits + Relevant Skill):} Reinforce wards, maintain circle, counter the entity's responses
\end{enumerate}

\textbf{Entity Resistance:}
During each phase, the possessing entity resists with a DV based on its Tier:
\begin{itemize}
\item \textbf{Tier I-II Entities:} DV 2
\item \textbf{Tier III-IV Entities:} DV 3
\item \textbf{Tier V+ Entities:} DV 4+
\end{itemize}

\textbf{Entity Response Actions:}
When the entity accumulates successes equal to or greater than the caster's total successes in a phase, it may attempt a response action that, if not resisted, advances the Exorcism Clock by one segment:
\begin{itemize}
\item \textbf{Whispers of Doubt (Social):} Tempt/corrupt a participant (Wits + Sway vs. DV 2-4)
\item \textbf{Physical Manifestation (Physical):} Cause harm to participants/subject (Wits + Melee vs. DV 2-4)
\item \textbf{Spiritual Assault (Mental):} Drain participant (Wits + Spirit vs. DV 2-4)
\item \textbf{Reality Distortion (Supernatural):} Create illusions/confusion (Wits + Arcana vs. DV 2-4)
\end{itemize}

All participant rolls to resist entity actions are made at Risky Position. Failure on resistance rolls generates 1 SB for the GM.

\textbf{Defensive Support:}
Non-participating allies may provide defense:
\begin{itemize}
\item Provide +1 die to one resistance roll per defender per phase
\item Defenders cannot take other actions during that phase
\end{itemize}

\textbf{Ritual Failure:}
If the Exorcism Clock fills completely before the ritual succeeds, the exorcism fails with severe consequences:
\begin{itemize}
\item The possessing entity grows stronger (+1 die to all future possession attempts by this entity)
\item The subject suffers additional corruption (advance Harm by one level or equivalent)
\item All participants suffer Fatigue 2 from spiritual exhaustion
\item The entity gains temporary influence over the location (start a Haunting clock [6])
\end{itemize}

\textbf{Ritual Success:}
If the caster achieves three consecutive phases where total successes exceed the entity's resistance DV by 2+, the entity is expelled:
\begin{itemize}
\item The possessing entity is banished from the subject
\item The subject stabilizes but requires extended care (Harm 2 recovery)
\item Caster and helpers suffer Fatigue 1 per Tier of the entity
\item Mark 1 XP for major spiritual conflict resolved
\end{itemize}

\textbf{For Player Characters:}
When a PC becomes possessed and exorcism fails, they gain ongoing complications:
\begin{itemize}
\item \textbf{Tier I-II Entities:} 1 SB banked each session from the entity's influence
\item \textbf{Tier III-IV Entities:} 2 SB banked each session from the entity's influence
\item \textbf{Tier V+ Entities:} 3 SB banked each session from the entity's influence
\end{itemize}

These banked SB represent the entity's ongoing influence and may manifest as:
\begin{itemize}
\item Compulsive behaviors or speech patterns
\item Nightmares or visions that provide the entity information
\item Uncontrolled moments where the entity acts through the character
\item Social consequences as others notice the character's changed nature
\end{itemize}

\shadow{Rukhir shivers, and I realize his feathers have gone dull. "Even I," he says quietly, "have felt the touch of things that would wear me like a glove. The difference, Veyla, is that I chose to be worn."}

\textbf{Player Agency Preservation:}
While possessed characters suffer ongoing complications, their players must always retain meaningful agency:
\begin{itemize}
\item Players choose when to resist the entity's influence
\item Major character decisions remain with the player
\item The entity cannot force the character to betray core relationships without resistance
\item Redemption paths should always remain possible, however difficult
\end{itemize}

The Ritual of Exorcism is not merely about casting spells—it is about the fundamental struggle between light and shadow, freedom and control, self and other. Approach it with the gravity it deserves, and remember that sometimes, the greatest mercy is to help a friend find peace, even if that peace requires them to let go of everything they were.

\textbf{Duration:} Variable encounter time, typically 10-15 minutes of game time per segment

\textbf{Casting Time:} Extended ritual requiring continuous focus and participation

\section*{Mist-Road Veil}
\textit{Draws a low fog that dulls footfalls and blurs silhouettes along a chosen path}

\textbf{Purpose:} Create concealing mist that aids stealthy movement along a specific route

\textbf{Primary Caster:} Wits 3+, Arcana 2+, Survival 2+

\textbf{Helpers:} 1-2 participants with Survival or Stealth skills

\textbf{Channel Phase (Wits + Survival):}
\begin{enumerate}
\item Gather water sources and cooling materials (damp cloth, morning dew, etc.)
\item Trace the path where mist is desired with light touch
\item Roll Wits + Survival for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Arcana):}
\begin{enumerate}
\item Channel moisture into the air along the designated path
\item Whisper words of concealment in the local tongue
\item Roll Wits + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} Mist is patchy or reveals more than it conceals; -1 die to Stealth rolls
\item \textbf{3-4 SB:} Mist attracts unwanted attention from creatures that hunt in fog; 2 SB added to next encounter
\item \textbf{5+ SB:} Mist becomes supernaturally thick, disorienting even the caster; party suffers -2 dice to navigation for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Creates concealing fog along a path up to 100 yards long for Significant Time
\item Party members moving along the path gain +2 dice to Stealth rolls
\item Sound is muffled by 50\% within the mist
\item Non-magical tracking becomes nearly impossible within the affected area
\end{itemize}

\textbf{Duration:} Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (10-15 minutes of preparation and ritual)

\section*{Bell of Truce}
\textit{Rings a hospitality note that compels all within earshot to parley before violence}

\textbf{Purpose:} Create a magical compulsion for peaceful negotiation before combat

\textbf{Primary Caster:} Presence 4+, Spirit 3+, Command 2+

\textbf{Helpers:} 1-2 participants with Sway or Performance skills

\textbf{Channel Phase (Spirit + Command):}
\begin{enumerate}
\item Obtain a bell or chime of significant quality
\item Focus on the concept of hospitality and peaceful resolution
\item Roll Spirit + Command for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Performance):}
\begin{enumerate}
\item Ring the bell with specific intent and rhythm
\item Speak words of truce in the old tongue
\item Roll Presence + Performance for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The compulsion is weak or easily resisted; -1 die to Sway rolls for next social encounter
\item \textbf{3-4 SB:} The bell's influence extends beyond intended area, affecting neutral parties; GM banks +2 SB for use in the next crowd-related encounter
\item \textbf{5+ SB:} The compulsion backfires, making the caster appear weak or cowardly; -2 dice to Command rolls for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item All intelligent beings within 200 yards must make Spirit + Resolve (DV 3) to initiate violence for Some Time
\item Those who successfully resist the compulsion gain +1 die to subsequent Sway rolls with the caster's party
\item Failure to parley first creates 2 SB backlash when violence finally erupts
\item The bell's tone marks a sacred space of temporary sanctuary
\end{itemize}

\textbf{Duration:} Some Time (approximately 30 minutes)

\textbf{Casting Time:} Some Time (ritual ringing and proclamation)

\section*{Lantern Weir}
\textit{Stitches light into a soft fence that guides crowds and resists panic surges}

\textbf{Purpose:} Create a barrier of light that guides movement and prevents chaotic rushes

\textbf{Primary Caster:} Wits 3+, Arcana 3+, Performance 2+

\textbf{Helpers:} 2-3 participants with Performance or Command skills

\textbf{Channel Phase (Wits + Performance):}
\begin{enumerate}
\item Gather multiple light sources (lanterns, candles, etc.)
\item Arrange participants in a line or curve to form the barrier
\item Roll Wits + Performance for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Weave the light sources together into a unified barrier
\item Speak words of guidance and protection
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The light barrier flickers or has gaps; -1 die to Crowd management rolls
\item \textbf{3-4 SB:} The barrier attracts attention from the wrong crowd; GM banks +2 SB for use in the next social encounter
\item \textbf{5+ SB:} The light becomes blinding rather than guiding; all within 50 yards suffer -1 die to perception rolls
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Creates a 50-foot barrier of soft, guiding light for Significant Time
\item Crowd movement is directed and calmed; +2 dice to Crowd management rolls
\item Panic effects are reduced by 2 SB for those within the barrier
\item The barrier provides dim illumination equivalent to moonlight
\end{itemize}

\textbf{Duration:} Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (15-20 minutes of arrangement and ritual)

\section*{Oath of Shared Breath}
\textit{Binds the party's rhythm; fatigue spreads thinly among all who consent}

\textbf{Purpose:} Distribute fatigue and minor harm among willing party members

\textbf{Primary Caster:} Presence 4+, Spirit 3+, Sway 2+

\textbf{Helpers:} All willing party members (2+ participants required)

\textbf{Channel Phase (Spirit + Sway):}
\begin{enumerate}
\item All participants join hands or touch in a circle
\item Each participant speaks their willingness to share the burden
\item Roll Spirit + Sway for primary caster, Spirit + Resolve for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Weave the participants' life forces together through ritual words
\item Create a shared rhythm of breath and heartbeat
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The sharing is uneven, favoring some over others; one participant suffers +1 Fatigue
\item \textbf{3-4 SB:} The bond creates unwanted emotional connections; party suffers -1 die to social rolls with outsiders for Significant Time
\item \textbf{5+ SB:} The sharing becomes too intimate; participants can sense each other's thoughts and emotions for one hour
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Creates a bond among willing participants for Significant Time
\item Total Fatigue among participants is divided equally (rounded up)
\item When one bonded participant would take Harm 1, the group may choose to distribute it: each participant takes 1 Fatigue instead
\item Participants can sense each other's general health (well/rested/wounded) and location within 100 yards
\end{itemize}

\textbf{Duration:} Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (10-15 minutes of ritual bonding)

\section*{The Bonding Ceremony}
\textit{A ritual to strengthen connections between companions}

\textbf{Purpose:} Create a permanent bond between party members that provides mechanical and narrative benefits

\textbf{Participants:} All willing party members (3-5 helpers maximum)

\textbf{Channel Phase (Wits + Sway):}
\begin{enumerate}
\item Each participant holds an object representing their connection to the group
\item Perform a shared activity (sharing a meal, telling stories, etc.) while focusing on the bond
\item Roll Wits + Sway skill for each participant
\end{enumerate}

\textbf{Weave Phase (Presence + Sway):}
\begin{enumerate}
\item Participants speak words of commitment to each other
\item Create a physical symbol of the bond (knotting cords, mixing earth from homelands, etc.)
\item Roll Presence + Sway for the primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} Temporary tension between bonded members (-1 die to cooperative actions for Significant Time)
\item \textbf{3-4 SB:} The bond creates unwanted attention from external forces; party generates 2 SB whenever they act as a group for the next session
\item \textbf{5+ SB:} All existing bonds between party members are temporarily severed; party suffers -2 dice to all social rolls for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Bonded members gain +1 die when aiding each other for one arc
\item Can sense each other's general emotional state across any distance for one arc
\item One "bond-mate" can be called upon for aid once per session (counts as Cap 3 follower for that action)
\end{itemize}

\textbf{Duration:} One full arc, renewable with simpler weekly reaffirmation

\textbf{Casting Time:} Significant Time (ritual requires shared meal and ceremony)

\section*{The Sanctuary Consecration}
\textit{A ritual to create a safe haven from supernatural threats}

\textbf{Purpose:} Establish a protected area that repels hostile supernatural entities

\textbf{Primary Caster:} Spirit 4+, Lore 3+

\textbf{Helpers:} 2-3 participants with relevant skills (Lore, Presence, Arcana)

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Gather materials sacred to the caster's beliefs or relevant to the threats faced
\item Consecrate the space through prayer, meditation, or ritual preparation
\item Roll Spirit + Lore for each participant
\end{enumerate}

\textbf{Weave Phase (Wits + Arcana):}
\begin{enumerate}
\item Perform the consecration ceremony within the designated area
\item Create permanent wards using blessed/sacred materials
\item Roll Wits + Arcana for the primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The sanctuary attracts the attention of other supernatural entities; 2 SB added to next supernatural encounter
\item \textbf{3-4 SB:} The protection creates a vulnerability elsewhere; choose one party member who becomes more susceptible to supernatural influence for one session
\item \textbf{5+ SB:} The ritual backfires, making the area supernaturally dangerous; the space becomes a beacon for hostile entities
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Area becomes sanctified ground (Sanctity +2) against specified supernatural threats for one month
\item Hostile entities of the specified type suffer -2 dice to actions within the area
\item Area provides +1 die to resistance rolls against specified supernatural effects for one month
\item Party members resting within the sanctuary recover 1 additional Fatigue per night
\end{itemize}

\textbf{Duration:} One month, renewable with weekly maintenance (10 minutes of ritual work)

\textbf{Casting Time:} Significant Time (requires preparation and ceremony)

\section*{The Memory Reconstruction}
\textit{A ritual to recover lost or suppressed memories}

\textbf{Purpose:} Retrieve information from a participant's suppressed or forgotten memories

\textbf{Primary Caster:} Wits 4+, Arcana 3+

\textbf{Helpers:} 1-2 participants with Insight or Lore skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Participant enters a deep meditative state
\item Primary caster prepares scrying tools or ritual components
\item Roll Wits + Arcana for primary caster, Wits + Insight/Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Spirit + Insight):}
\begin{enumerate}
\item Guide the participant through their memories using guided visualization
\item Use ritual tools to stabilize recovered memories
\item Roll Spirit + Insight for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} Recovered memories are fragmented or unreliable; information may be misleading
\item \textbf{3-4 SB:} The ritual triggers traumatic memories causing temporary psychological effects; participant suffers -1 die to social rolls for Significant Time
\item \textbf{5+ SB:} Participant becomes obsessed with recovered memories; must make Spirit + Resolve (DV 2) rolls to act against the memory's implications for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Recover one specific memory or set of memories from the participant's past (immediate)
\item Memory clarity depends on success margin (partial recall vs. complete recovery)
\item Participant gains +2 dice on related knowledge rolls for this scene
\item GM reveals one piece of information that connects the recovered memory to current events
\end{itemize}

\textbf{Duration:} Memory effects last for this scene; knowledge benefits last for this session

\textbf{Casting Time:} Some Time (requires deep meditation and guided ritual)

\section*{The Shadow Passage}
\textit{A ritual to travel through the Ways Between}

\textbf{Purpose:} Transport the party across great distances through shadow-realm shortcuts

\textbf{Primary Caster:} Wits 5+, Arcana 4+, Spirit 3+

\textbf{Helpers:} All party members (spiritual support during the journey)

\textbf{Channel Phase (Spirit + Arcana):}
\begin{enumerate}
\item Create a circle of shadow-conductive materials (obsidian, black candles, etc.)
\item Each participant contributes a drop of blood or lock of hair to the circle
\item Roll Spirit + Arcana for primary caster, Spirit + Resolve for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Arcana):}
\begin{enumerate}
\item Perform the opening ceremony to breach the shadow-realm
\item Lead the party through the passage while maintaining the connection
\item Roll Wits + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The journey takes longer than expected, arriving at an inconvenient time (advance relevant clocks by 1 segment)
\item \textbf{3-4 SB:} The party arrives in a slightly different location than intended; position for first action is Risky instead of Controlled
\item \textbf{5-6 SB:} Something follows the party back from the shadow-realm; introduce a persistent supernatural threat that tracks the party
\item \textbf{7+ SB:} The party finds themselves in the Ways Between.
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Transport party instantly across great distances (anywhere on the same continent) for this scene
\item Arrival position is Controlled for planned destinations, Risky for improvised ones
\item +1 effect on the first action after arrival (momentum from the journey)
\item Party is immune to fatigue from the travel itself
\item Cannot be used more than once per week without increasing backlash severity
\end{itemize}

\textbf{Duration:} Instantaneous travel with lingering effects for this scene

\textbf{Casting Time:} Significant Time (complex ritual requiring careful preparation)

\section*{The Geas of Binding}
\textit{A ritual to create unbreakable oaths or compulsions}

\textbf{Purpose:} Bind a willing or unwilling target to a specific course of action

\textbf{Primary Caster:} Presence 4+, Spirit 4+, Lore 3+

\textbf{Helpers:} 1-2 participants with relevant cultural or religious knowledge

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Prepare the binding materials according to cultural tradition
\item Research the target's nature and vulnerabilities
\item Roll Spirit + Lore for primary caster, relevant skill for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Perform the binding ceremony with the target present
\item Speak the terms of the geas in the appropriate ritual language
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The geas has unintended loopholes or consequences; target finds creative ways to technically fulfill the geas while undermining its intent
\item \textbf{3-4 SB:} The caster becomes partially bound by the same geas; must also follow a related constraint for Significant Time
\item \textbf{5+ SB:} The binding creates a spiritual debt that attracts unwanted supernatural attention; 3 SB added to next supernatural encounter involving the caster
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Target is compelled to follow the terms of the geas or suffer spiritual consequences for one arc
\item Breaking the geas requires a successful Spirit + Resolve roll (DV 4+) each time
\item Failure to follow the geas causes 1 SB per day of violation
\item Geas can be complex but must be clearly defined (no "be good" - specify actions)
\item Willing targets suffer reduced backlash; unwilling targets may resist with opposed roll
\end{itemize}

\textbf{Duration:} One full arc, permanent unless specifically countered by equal magic

\textbf{Casting Time:} Significant Time (formal ceremony requiring proper ritual components)

\section*{The Communion of the Ancients}
\textit{A ritual to speak with the spirits of the dead}

\textbf{Purpose:} Contact and communicate with deceased individuals or ancestral spirits

\textbf{Primary Caster:} Spirit 4+, Lore 3+, Presence 2+

\textbf{Helpers:} 1-2 participants with spiritual sensitivity or relevant cultural knowledge

\textbf{Channel Phase (Spirit + Presence):}
\begin{enumerate}
\item Prepare the ritual space according to cultural traditions for contacting the dead
\item Offer appropriate sacrifices or prayers to the spirits
\item Roll Spirit + Presence for primary caster, Spirit + Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Lore):}
\begin{enumerate}
\item Open the channel to the spirit realm
\item Guide the summoned spirit toward communication
\item Roll Wits + Lore for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The contacted spirit is confused or unhelpful; information provided is vague or misleading
\item \textbf{3-4 SB:} Other, less welcome spirits are drawn to the ritual; 2 SB added to next supernatural encounter
\item \textbf{5+ SB:} The ritual creates a permanent spiritual connection that cannot be easily severed; caster suffers -1 die to all social rolls involving living beings for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Communicate with one specific deceased individual or type of ancestral spirit for Some Time
\item Spirit can provide information, advice, or warnings about specified topics
\item Spirit's knowledge is limited by their mortal understanding and perspective
\item Communication lasts for a number of minutes equal to the success margin
\item Spirit may request a favor or service in return for their cooperation
\end{itemize}

\textbf{Duration:} Some Time (ritual communication period)

\textbf{Casting Time:} Significant Time (preparation and ceremony required)

\section*{The Ward of Protection}
\textit{A ritual to create lasting magical defenses}

\textbf{Purpose:} Establish permanent or semi-permanent protective wards

\textbf{Primary Caster:} Wits 4+, Arcana 3+, relevant cultural skill

\textbf{Helpers:} 2-3 participants with Craft, Engineering, or cultural ritual skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Gather materials appropriate to the type of protection desired
\item Prepare the location where wards will be placed
\item Roll Wits + Arcana for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Craft):}
\begin{enumerate}
\item Inscribe, construct, or otherwise create the physical components of the wards
\item Activate the wards through ritual consecration
\item Roll Wits + Craft for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The wards have a minor vulnerability or require specific activation conditions; wards function at -1 effectiveness until properly calibrated
\item \textbf{3-4 SB:} The wards attract the attention of entities they're designed to repel; 2 SB added to next encounter with specified threat type
\item \textbf{5+ SB:} The wards backfire, creating the opposite effect or affecting unintended targets; wards harm allies instead of protecting them for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create wards that provide +2 dice to defense against specified threats for one month
\item Wards can cover an area, protect specific individuals, or guard objects
\item Duration: One month, renewable with monthly maintenance rituals
\item Wards can be complex (multiple effects) but each additional effect increases backlash risk
\item Wards can be detected by those with appropriate skills and may reveal the caster's presence
\end{itemize}

\textbf{Duration:} One month, renewable with weekly maintenance

\textbf{Casting Time:} Significant Time (complex construction and consecration process)

\section*{The Veil of Misdirection}
\textit{A ritual to conceal the party's true nature and intentions}

\textbf{Purpose:} Create illusions and false impressions that mask the party's activities

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Deception 2+

\textbf{Helpers:} 2-3 participants with Performance, Deception, or cultural masking skills

\textbf{Channel Phase (Wits + Deception):}
\begin{enumerate}
\item Prepare illusion components (cosmetics, costumes, symbolic objects)
\item Each participant contributes a personal item that represents their true self
\item Roll Wits + Deception for primary caster, relevant skills for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Performance):}
\begin{enumerate}
\item Perform the masquerade ritual, weaving false identities around each participant
\item Create shared illusion patterns that reinforce the group's fabricated story
\item Roll Presence + Performance for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The illusions have minor flaws or inconsistencies; -1 die to social rolls when under scrutiny
\item \textbf{3-4 SB:} The false identities begin to influence the participants' behavior; party members must make Wits + Resolve (DV 2) rolls to act against their assumed personas for Significant Time
\item \textbf{5+ SB:} The ritual creates a feedback loop where the false identities become partially real; participants suffer identity confusion and -2 dice to rolls involving their true skills for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Party assumes completely fabricated identities that are believable to casual observers for one session
\item Each participant gains +2 dice to Deception and Performance rolls related to their assumed identity
\item Social interactions start Controlled when the false identities are relevant
\item Magical detection of true identities suffers -2 dice for Significant Time
\item Party can coordinate complex deceptions as a group action with +1 effect
\end{itemize}

\textbf{Duration:} One full session, requiring daily renewal (10 minutes of ritual work)

\textbf{Casting Time:} Significant Time (elaborate preparation and group ceremony)

\section*{The Art of Listening}

In the Tulkani tradition, we say that the greatest magic is not in speaking, but in hearing what others dare not say. The rituals in this chapter will teach you to listen to the whispers that exist in the spaces between words, to read the truth in a person's eyes, and to gather secrets that would make emperors tremble.

\section*{The Listening Shadow}
\textit{A ritual to hear distant conversations}

\textbf{Purpose:} To hear conversations from great distances, particularly those that the speakers believe are private

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Insight 2+

\textbf{Helpers:} 1 participant with Insight skill

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Light a black candle at the darkest hour of night, when even the moon has set
\item Take a raven's feather and whisper into it the name of the person whose words you wish to hear
\item Bury the feather beneath your doorstep, or in the earth beneath a window where your target might pass
\item Roll Wits + Arcana for primary caster, Wits + Insight for helper
\end{enumerate}

\textbf{Weave Phase (Spirit + Insight):}
\begin{enumerate}
\item For three nights, you must dream of shadows dancing. Do not fight these dreams - embrace them
\item On the fourth night, perform the final scrying
\item Roll Spirit + Insight for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} You hear fragments of many conversations, making it difficult to isolate the target; -1 die to subsequent Investigation rolls for one scene
\item \textbf{3-4 SB:} You hear one conversation that you desperately wish not to hear; GM introduces a disturbing piece of knowledge that burdens the caster
\item \textbf{5+ SB:} The target becomes aware they are being spied upon; they gain +2 dice against the caster in all future social interactions
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Hear a specific conversation involving your target for Significant Time
\item Gain the ability to distinguish truth from lies in any spoken word for the duration, but only while holding your breath
\item Information gathered provides +2 dice to related Investigation or Social rolls for this session
\end{itemize}

\textbf{Duration:} Significant Time (approximately 1 hour of eavesdropping)

\textbf{Casting Time:} Extended (requires 4 nights of preparation, final casting on fourth night)

\textbf{Price:} You will hear one conversation that you desperately wish not to hear. This knowledge will burden you until you either act upon it or deliberately forget it.

\textbf{Warning:} Do not use this ritual to spy upon those who are innocent of wrongdoing. The Patron favors those who seek justice, not those who indulge in cruelty.

\section*{The Cleansing Fire}
\textit{A ritual to purify corrupted places and objects}

\textbf{Purpose:} Remove curses, malevolent enchantments, or spiritual contamination

\textbf{Primary Caster:} Spirit 4+, Arcana 3+, Lore 2+

\textbf{Helpers:} 2-3 participants with Spirit or relevant cultural skills

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Gather purifying materials (sacred herbs, blessed water, salt, etc.)
\item Research the nature of the corruption to be cleansed
\item Roll Spirit + Lore for primary caster, Spirit + relevant skill for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Perform the cleansing ceremony with fire as the central element
\item Speak words of purification in the appropriate ritual language
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The cleansing is incomplete; the corruption is reduced but not eliminated, returning at half strength in one day
\item \textbf{3-4 SB:} The ritual agitates the corruption, causing it to manifest more aggressively before it can be cleansed; 2 SB added to next encounter with the corruption
\item \textbf{5+ SB:} The cleansing fire turns inward, affecting the caster instead; caster suffers Harm 1 and gains a new minor curse related to fire or purification
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Remove one curse, enchantment, or spiritual contamination from a person, object, or area
\item Affected area or object becomes spiritually clean for one month
\item Beneficiaries gain +1 die to resistance rolls against similar corruptions for Significant Time
\item If used on a possessed person, the entity suffers -1 die to all actions for the rest of the scene
\end{itemize}

\textbf{Duration:} Immediate effect with lingering benefits for one month

\textbf{Casting Time:} Significant Time (requires preparation and ceremony)

\section*{The Dreaming Stone}
\textit{A ritual to send messages through dreams}

\textbf{Purpose:} Communicate with a specific person through their dreams

\textbf{Primary Caster:} Spirit 4+, Arcana 3+, Insight 2+

\textbf{Helpers:} 1 participant with Insight or Lore skill

\textbf{Channel Phase (Spirit + Insight):}
\begin{enumerate}
\item Obtain a personal item from the target or something they have touched recently
\item Prepare the dream-sending stone or crystal
\item Roll Spirit + Insight for primary caster, Wits + Insight/Lore for helper
\end{enumerate}

\textbf{Weave Phase (Wits + Arcana):}
\begin{enumerate}
\item Focus on the target's essence through the personal item
\item Shape the dream-message using symbolic imagery
\item Roll Wits + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The message is distorted or misinterpreted; -1 die to related social rolls when the message is delivered in waking life
\item \textbf{3-4 SB:} The target's subconscious fights back, invading the caster's dreams; caster suffers Fatigue 1 and vivid nightmares for Significant Time
\item \textbf{5+ SB:} The dream-bridge becomes two-way; the target can now send dream-messages to the caster or even attempt to possess them in dreams
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Send a specific message or image to the target's dreams for one night
\item Message clarity depends on success margin (symbolic hints vs. clear communication)
\item Target gains +2 dice to related rolls if the dream-message provides useful information or warning
\item Can be used to plant suggestions (with target's resistance roll) or provide comfort/encouragement
\end{itemize}

\textbf{Duration:} One night (message delivered during target's next sleep)

\textbf{Casting Time:} Some Time (requires meditation and focus)

\section*{The Forge of Names}
\textit{A ritual to create or bind true names}

\textbf{Purpose:} Discover, create, or bind true names for magical working

\textbf{Primary Caster:} Wits 5+, Lore 4+, Arcana 3+

\textbf{Helpers:} 1-2 participants with Lore or cultural naming knowledge

\textbf{Channel Phase (Wits + Lore):}
\begin{enumerate}
\item Gather materials representing the subject (hair, blood, significant possession, or symbolic representation)
\item Research the subject's history, connections, and essence
\item Roll Wits + Lore for primary caster, Wits + Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Spirit + Arcana):}
\begin{enumerate}
\item Perform the naming ceremony, speaking the syllables as they are revealed
\item Bind the true name to a physical token or within the caster's memory
\item Roll Spirit + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The name is incomplete or has a minor flaw; using it provides only half the normal benefits
\item \textbf{3-4 SB:} The subject becomes aware their true name has been discovered; they gain +1 die to resist any magic using their name for one arc
\item \textbf{5+ SB:} The naming ritual backfires, revealing the caster's own true name to the subject or to malevolent forces listening
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Discover or create the true name of a person, place, or thing
\item True name can be used for various magical effects (control, banishment, enhanced understanding)
\item Having a true name provides +2 dice to any magic specifically targeting that subject
\item Subject suffers -1 die to resist magic from the caster who knows their true name
\item Name can be bound to a token, making it usable by others who possess the token
\end{itemize}

\textbf{Duration:} Permanent (true name is forever linked to the subject)

\textbf{Casting Time:} Significant Time (extensive research and ceremony)

\section*{The Mirror's Gaze}
\textit{A ritual to see through the eyes of others}

\textbf{Purpose:} Temporarily perceive through the senses of a willing or unwitting subject

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Insight 2+

\textbf{Helpers:} 1 participant with Insight skill

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Prepare a reflective surface (mirror, pool of water, polished metal)
\item Establish connection to the target through a personal item or direct line of sight
\item Roll Wits + Arcana for primary caster, Wits + Insight for helper
\end{enumerate}

\textbf{Weave Phase (Spirit + Insight):}
\begin{enumerate}
\item Gaze into the reflective surface while focusing on the target
\item Merge consciousness with the target's senses
\item Roll Spirit + Insight for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The connection is unstable; vision is blurred or intermittent for the duration
\item \textbf{3-4 SB:} The target senses the intrusion; if unwilling, they gain +2 dice to any rolls to detect or resist the caster's presence for one session
\item \textbf{5+ SB:} Consciousness becomes trapped in the target's mind; caster's body falls into a coma while spirit remains lost in the target's perception
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item See through the eyes of the target for Some Time
\item Gain access to the target's immediate sensory input (sight, hearing, etc.)
\item If target is willing, gain +1 die to related Investigation or Stealth rolls using their perspective
\item Can be used to guide a lost target or scout dangerous areas remotely
\end{itemize}

\textbf{Duration:} Some Time (approximately 30 minutes of shared perception)

\textbf{Casting Time:} Some Time (requires focus and preparation)

\section*{The Bloodbound Pact}
\textit{A ritual to create magical agreements enforced by mutual consequence}

\textbf{Purpose:} Create a binding agreement where breaking the pact harms all parties involved

\textbf{Primary Caster:} Presence 4+, Spirit 4+, Lore 3+

\textbf{Helpers:} All parties involved in the pact (minimum 2 parties)

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Each party must contribute a drop of blood to the pact vessel
\item Define the exact terms of the agreement in precise language
\item Roll Spirit + Lore for primary caster, Spirit + Resolve for each participating party
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Speak the terms of the pact in the old tongue while mixing the blood
\item Each party must swear the oath and accept the consequences
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The pact has minor loopholes or ambiguities; parties may find legalistic ways to avoid true compliance
\item \textbf{3-4 SB:} The magical enforcement is stronger than intended; even minor infractions trigger harsh consequences
\item \textbf{5+ SB:} The pact becomes contagious, drawing unrelated parties into its terms or consequences
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create a magically binding agreement between all participating parties
\item If any party breaks the pact, all parties suffer Harm 1 per severity level of the breach
\item Pact remains active for one arc or until fulfilled/voided through specific counter-ritual
\item Parties gain +1 die to cooperative actions related to fulfilling the pact terms
\item Breaking the pact creates 2 SB that can be spent by the GM to complicate the breaker's future actions
\end{itemize}

\textbf{Duration:} One full arc, or until pact is completed/voided

\textbf{Casting Time:} Significant Time (formal ceremony requiring all parties)

\section*{The Stone's Memory}
\textit{A ritual to read the history imprinted upon objects or locations}

\textbf{Purpose:} Perceive significant events associated with an object or location

\textbf{Primary Caster:} Spirit 4+, Insight 3+, Lore 2+

\textbf{Helpers:} 1-2 participants with Insight or Investigation skills

\textbf{Channel Phase (Spirit + Insight):}
\begin{enumerate}
\item Hold or place hands upon the object/location to be read
\item Clear the mind and open the spirit to receive impressions
\item Roll Spirit + Insight for primary caster, Wits + Insight/Investigation for helpers
\end{enumerate}

\textbf{Weave Phase (Wits + Lore):}
\begin{enumerate}
\item Focus on specific questions or time periods if desired
\item Allow the memories to flow through consciousness
\item Roll Wits + Lore for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The memories are fragmented or confusing; -1 die to subsequent Investigation rolls for one scene
\item \textbf{3-4 SB:} The caster becomes overwhelmed by traumatic memories; suffers Fatigue 1 and -1 die to social rolls for Significant Time
\item \textbf{5+ SB:} The caster becomes trapped in a memory loop, reliving the object's history; requires external intervention to break the connection
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Witness 1-3 significant events associated with the object/location (based on success margin)
\item Events are experienced as vivid sensory impressions rather than clear narrative
\item Gain +2 dice to related Investigation or Lore rolls for this session
\item Can detect if the object/location is currently under magical influence
\end{itemize}

\textbf{Duration:} Some Time (approximately 20-30 minutes of memory viewing)

\textbf{Casting Time:} Some Time (requires meditation and focus)

\section*{The Wind's Whispers}
\textit{A ritual to gather information from the environment}

\textbf{Purpose:} Collect news, rumors, and information carried by natural elements

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Investigation 2+

\textbf{Helpers:} 1-2 participants with Investigation or Survival skills

\textbf{Channel Phase (Wits + Investigation):}
\begin{enumerate}
\item Position yourself in an area where information might naturally gather (crossroads, market,风口)
\item Prepare listening tools (horn, focusing crystal, etc.) if available
\item Roll Wits + Investigation for primary caster, Wits + Investigation/Survival for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Cast the spell during a time of natural transition (dawn, dusk, changing weather)
\item Speak words to call the information-bearing winds
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The information gathered is mostly gossip or irrelevant; -1 die to subsequent Investigation rolls for one scene
\item \textbf{3-4 SB:} The ritual attracts unwanted attention from those who prefer their secrets stay hidden; 2 SB added to next social encounter involving the caster
\item \textbf{5+ SB:} The caster becomes overwhelmed by the volume of information; suffers Fatigue 1 and -2 dice to all mental actions for Significant Time
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Gather recent news, rumors, and information from the surrounding area for the past week
\item Information is received as fragmented whispers, requiring interpretation
\item Gain +2 dice to related Social or Investigation rolls for this session
\item Can specifically target information about one subject at the cost of overall breadth
\end{itemize}

\textbf{Duration:} Information gathering occurs over Significant Time (approximately 1 hour)

\textbf{Casting Time:} Some Time (requires positioning and brief ceremony)

\section*{The Heart's True Measure}
\textit{A ritual to determine the alignment and intentions of others}

\textbf{Purpose:} Assess the moral character and current intentions of a person or group

\textbf{Primary Caster:} Spirit 4+, Insight 3+, Presence 2+

\textbf{Helpers:} 1 participant with Insight or Sway skill

\textbf{Channel Phase (Spirit + Insight):}
\begin{enumerate}
\item Observe the target(s) from a distance for at least one hour
\item Prepare ritual materials (candle, incense, focus crystal)
\item Roll Spirit + Insight for primary caster, Wits + Insight/Sway for helper
\end{enumerate}

\textbf{Weave Phase (Presence + Insight):}
\begin{enumerate}
\item Perform the assessment ritual while maintaining focus on the target(s)
\item Interpret the signs and omens that reveal true nature
\item Roll Presence + Insight for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The assessment is partially clouded by the target's own self-deception; -1 accuracy in reading their current intentions
\item \textbf{3-4 SB:} The target becomes aware they are being scrutinized; they gain +1 die to resist future social manipulation by the caster for one session
\item \textbf{5+ SB:} The ritual reveals uncomfortable truths about the caster's own nature; -2 dice to social rolls for one session as inner conflicts surface
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Determine the general moral alignment of the target(s) (Lawful/Chaotic, Good/Evil, or Neutral variants)
\item Gain insight into the target's current intentions and motivations
\item Receive a sense of whether the target is lying or concealing information
\item Gain +1 die to all Social rolls against the target(s) for this session based on understanding their nature
\end{itemize}

\textbf{Duration:} Assessment is valid for this session

\textbf{Casting Time:} Significant Time (requires extended observation and ceremony)

\section*{The Chain That Binds}
\textit{A ritual to create sympathetic connections for remote influence}

\textbf{Purpose:} Establish a magical link to a person or object for remote sensing or minor influence

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Lore 2+

\textbf{Helpers:} 1 participant with Lore or Tinker skill

\textbf{Channel Phase (Wits + Lore):}
\begin{enumerate}
\item Obtain a significant personal item from the target or a piece of the target itself
\item Prepare linking materials (chain, cord, matched objects)
\item Roll Wits + Lore for primary caster, Wits + Lore/Tinker for helper
\end{enumerate}

\textbf{Weave Phase (Spirit + Arcana):}
\begin{enumerate}
\item Perform the binding ceremony, linking the target item to the caster's focus
\item Seal the connection with a drop of the caster's blood or essence
\item Roll Spirit + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The connection is weak or intermittent; -1 die to all effects using the link for one session
\item \textbf{3-4 SB:} The target becomes aware of the connection; they gain +1 die to resist any influence attempts through the link for one arc
\item \textbf{5+ SB:} The connection becomes parasitic, draining the caster's energy; suffer Fatigue 1 per day the link remains active
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create a sympathetic link to the target for one arc
\item Sense the target's general health and emotional state across any distance
\item Inflict 1 Fatigue on the target to restore 1 Fatigue to the caster (once per day)
\item Gain +1 die to any divination magic specifically targeting the linked subject
\item Can be used to send simple sensations or warnings to the target (if willing)
\end{itemize}

\textbf{Duration:} One arc, renewable with daily 5-minute maintenance ritual

\textbf{Casting Time:} Some Time (requires careful preparation and ceremony)

\section*{The Veil of Years}
\textit{A ritual to obscure or age objects and appearances}

\textbf{Purpose:} Alter the apparent age of objects or beings, or hide their true nature

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Tinker 2+

\textbf{Helpers:} 1-2 participants with Tinker or Craft skills

\textbf{Channel Phase (Wits + Tinker):}
\begin{enumerate}
\item Examine the object or person to be altered in detail
\item Gather materials appropriate to the desired aging effect (dust, patina agents, temporal focus)
\item Roll Wits + Tinker for primary caster, Wits + Tinker/Craft for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Apply the aging/veiling effect through touch and ritual words
\item Shape the temporal illusion to achieve the desired appearance
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The effect is superficial or temporary; ages normally within one day
\item \textbf{3-4 SB:} The aging process affects the caster as well; suffer 1 Fatigue and appear slightly older for one session
\item \textbf{5+ SB:} The temporal manipulation goes awry; the target ages rapidly (or de-ages) one year for every hour the spell was meant to last
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Age or de-age an object by 10-100 years per success level
\item Alter a person's apparent age by up to 20 years (if willing)
\item Hide magical auras or true nature of objects from casual inspection
\item Gain +2 dice to Deception rolls involving the altered object/person for this scene
\item Effect can be made permanent with additional ritual work
\end{itemize}

\textbf{Duration:} Scene (unless made permanent through extended ritual)

\textbf{Casting Time:} Some Time (15-20 minutes of careful work)

\section*{The Song of Shifting Tides}
\textit{A ritual to influence luck and probability in small ways}

\textbf{Purpose:} Bend chance and probability to favor the caster's intentions

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Performance 2+

\textbf{Helpers:} 2-3 participants with Performance or Luck-based skills

\textbf{Channel Phase (Wits + Performance):}
\begin{enumerate}
\item Begin a performance or chant that resonates with the desired outcome
\item Focus the group's intent on the specific area of luck to be influenced
\item Roll Wits + Performance for primary caster, Presence + Performance for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Weave the performance into a magical pattern that influences probability
\item Direct the flow of chance toward favorable outcomes
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The luck evens out; an equally beneficial coincidence occurs for an enemy or不利 situation arises for an ally
\item \textbf{3-4 SB:} The probability manipulation creates a paradox; -1 die to all rolls involving chance for Significant Time
\item \textbf{5+ SB:} Fortune turns against the caster; all rolls involving chance suffer -2 dice for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Influence minor probability for one scene (re-roll 1s on all dice pools for the party)
\item Grant one significant stroke of luck to a willing party member (GM discretion)
\item Cause one minor mishap or inconvenience to an enemy or obstacle
\item Gain +1 effect on one action that relies on timing or favorable circumstances
\end{itemize}

\textbf{Duration:} One scene (approximately 30 minutes of influenced events)

\textbf{Casting Time:} Some Time (requires coordinated performance)

\section*{The Anchor of Souls}
\textit{A ritual to protect against death and spiritual harm}

\textbf{Purpose:} Create a magical safeguard that can return a soul from the brink of death

\textbf{Primary Caster:} Spirit 5+, Lore 4+, Presence 3+

\textbf{Helpers:} 2-3 participants with Spirit or religious/cultural skills

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Prepare the soul-anchor (precious stone, blessed metal, or sacred relic)
\item Perform the consecration rites specific to the caster's tradition
\item Roll Spirit + Lore for primary caster, Spirit + relevant skill for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Bind the anchor to the intended recipient through ritual words and touch
\item Establish the conditions under which the anchor will activate
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The anchor's activation conditions are too broad or narrow; it may activate when not needed or fail when needed most
\item \textbf{3-4 SB:} The anchor draws attention from entities that feed on near-death experiences; 2 SB added to next supernatural encounter
\item \textbf{5+ SB:} The anchor becomes cursed, drawing the recipient toward dangerous situations to test its power; -1 die to rolls to avoid obviously perilous actions
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create a soul-anchor that activates when the recipient would otherwise die
\item Upon activation, the recipient is stabilized at 1 Health point and teleported to the anchor's location
\item Anchor remains active for one month or until used
\item Recipient gains +1 die to all resistance rolls against death effects while anchor is active
\item Anchor can be set to protect against specific types of death (magical, physical, spiritual)
\end{itemize}

\textbf{Duration:} One month or until activated

\textbf{Casting Time:} Significant Time (elaborate ceremony requiring rare materials)

\section*{The Web of Whispers}
\textit{A ritual to create a network of informants and information gathering}

\textbf{Purpose:} Establish a magical network for gathering information across a region

\textbf{Primary Caster:} Wits 4+, Arcana 3+, Sway 3+

\textbf{Helpers:} 3-5 participants with Social or Investigation skills

\textbf{Channel Phase (Wits + Sway):}
\begin{enumerate}
\item Recruit willing participants to serve as nodes in the information network
\item Provide each participant with a token or signal for communication
\item Roll Wits + Sway for primary caster, Presence + Sway/Investigation for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Link the participants together through a shared magical pattern
\item Establish protocols for information gathering and reporting
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The network is inefficient or prone to misinformation; -1 die to all information gathering through the network for one week
\item \textbf{3-4 SB:} The network attracts unwanted attention from authorities or rival organizations; 2 SB added to next social encounter involving information trading
\item \textbf{5+ SB:} The network becomes compromised, with enemies infiltrating as false nodes; all information from the network is suspect for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create an information network spanning a significant geographical area
\item Gain access to local news, rumors, and gossip from all network nodes
\item Can request specific information about topics or individuals within the network's range
\item Network provides +2 dice to Investigation or Social rolls involving information gathering for one month
\item One participant per week can be called upon for a significant favor related to information
\end{itemize}

\textbf{Duration:} One month, renewable with weekly 15-minute maintenance ritual

\textbf{Casting Time:} Significant Time (requires recruitment and coordination)

\section*{The Forge of Elements}
\textit{A ritual to transmute base materials into other substances}

\textbf{Purpose:} Transform one type of material into another through magical means

\textbf{Primary Caster:} Wits 5+, Arcana 4+, Tinker 3+

\textbf{Helpers:} 2-3 participants with Tinker, Craft, or Engineering skills

\textbf{Channel Phase (Wits + Tinker):}
\begin{enumerate}
\item Gather the base materials to be transformed and the catalyst substances
\item Prepare the transmutation array or forge according to elemental principles
\item Roll Wits + Tinker for primary caster, Wits + Tinker/Craft for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Arcana):}
\begin{enumerate}
\item Apply magical energy to reshape the material's fundamental structure
\item Speak the words of transformation in the language of creation
\item Roll Presence + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The transmutation is imperfect; the resulting material has flaws or reduced quality
\item \textbf{3-4 SB:} The transformation process is unstable; 2 SB added to next crafting or tinkering action involving the caster
\item \textbf{5+ SB:} The transmutation triggers a chain reaction, affecting unintended materials in the vicinity; area suffers 1 segment of environmental damage clock
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Transform up to 10 pounds of base material into another substance of equal or lesser value per success level
\item Can create alloys, treat materials for specific properties, or purify substances
\item Gain +2 dice to subsequent Tinker or Craft rolls involving the transmuted materials
\item Complex transmutations (precious metals, magical materials) require higher success margins and increase backlash risk
\item Effect can be made permanent with proper sealing rituals
\end{itemize}

\textbf{Duration:} Permanent (unless reversed by specific counter-ritual)

\textbf{Casting Time:} Significant Time (requires specialized setup and careful procedure)

\section*{The Mirror of Fates}
\textit{A ritual to glimpse possible futures and their consequences}

\textbf{Purpose:} Peer into potential future outcomes based on current decisions

\textbf{Primary Caster:} Wits 5+, Arcana 4+, Insight 3+

\textbf{Helpers:} 1-2 participants with Insight or Lore skills

\textbf{Channel Phase (Wits + Insight):}
\begin{enumerate}
\item Define the specific decision or action to be examined
\item Prepare the scrying medium (mirror, pool, crystal ball)
\item Roll Wits + Insight for primary caster, Wits + Insight/Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Spirit + Arcana):}
\begin{enumerate}
\item Focus magical energy through the scrying medium to pierce the veil of time
\item Observe the branching paths and their potential outcomes
\item Roll Spirit + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The visions are cloudy or metaphorical; -1 die to subsequent Investigation rolls related to future planning for one scene
\item \textbf{3-4 SB:} The caster becomes fixated on one possible future; -1 die to rolls involving alternative approaches for Significant Time
\item \textbf{5+ SB:} The ritual reveals a catastrophic future that cannot be easily forgotten; caster suffers Fatigue 1 and -2 dice to all rolls for one session due to despair or obsession
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Glimpse 1-3 possible futures stemming from a specific decision or action
\item Futures are shown with varying degrees of clarity based on success margin
\item Gain +2 dice to rolls involving planning or preparation based on the visions for this session
\item Can specifically request to see consequences of one particular course of action
\item GM may reveal one crucial detail that changes the interpretation of all futures shown
\end{itemize}

\textbf{Duration:} Visions last for this scene; knowledge benefits last for this session

\textbf{Casting Time:} Some Time (requires deep concentration and scrying preparation)

\section*{The Bond of Life}
\textit{A ritual to share life force between willing participants}

\textbf{Purpose:} Transfer health and vitality between bonded individuals

\textbf{Primary Caster:} Spirit 4+, Presence 3+, Lore 2+

\textbf{Helpers:} All willing participants (minimum 2, maximum 6)

\textbf{Channel Phase (Spirit + Presence):}
\begin{enumerate}
\item All participants join hands or touch in a circle
\item Each participant must willingly offer a portion of their life force
\item Roll Spirit + Presence for primary caster, Spirit + Resolve for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Weave the life forces together through ritual words and shared breath
\item Establish the conditions under which the bond will activate
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The life force sharing is inefficient; more energy is lost in the transfer than gained by the recipient
\item \textbf{3-4 SB:} The bond creates unwanted empathy; participants suffer -1 die to rolls involving personal desires that conflict with the group's needs for Significant Time
\item \textbf{5+ SB:} The life force exchange becomes uncontrolled; all participants suffer Harm 1 as energy flows chaotically between them
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Create a life-bond between all participating members for one session
\item Any participant can transfer 1 Health point to another bonded member as a reflexive action
\item If one bonded member would die, they can draw 1 Health point from each other bonded member (maximum transfer equal to number of other participants)
\item All participants gain +1 die to resistance rolls against death or life-draining effects while the bond is active
\item Participants can sense each other's general health and location within 500 yards
\end{itemize}

\textbf{Duration:} One session, renewable with 10-minute daily reaffirmation ritual

\textbf{Casting Time:} Some Time (requires circle formation and shared ceremony)

\section*{The Key of Locked Gates}
\textit{A ritual to unlock magical seals and bypass barriers}

\textbf{Purpose:} Open magically sealed doors, containers, or passages

\textbf{Primary Caster:} Wits 4+, Arcana 4+, Tinker 2+

\textbf{Helpers:} 1-2 participants with Tinker or Lore skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Examine the sealed object or barrier to understand its nature
\item Gather tools and components appropriate to the type of seal
\item Roll Wits + Arcana for primary caster, Wits + Tinker/Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Tinker):}
\begin{enumerate}
\item Apply magical and physical techniques to unravel the seal
\item Speak words of opening in the language appropriate to the barrier's creation
\item Roll Presence + Tinker for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The seal is only partially opened or will relock quickly; -1 die to subsequent attempts to keep the passage open
\item \textbf{3-4 SB:} The unlocking triggers security measures or alarms; 2 SB added to next encounter in the area
\item \textbf{5+ SB:} The seal was a containment ward; whatever it was holding back begins to escape
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Unlock or bypass one magical seal, ward, or barrier
\item Effect strength based on success margin (simple locks vs. complex wards)
\item Gain +2 dice to subsequent Tinker or Arcana rolls involving lock/ward analysis for this session
\item Can create a temporary key or passcode that works for one day
\item Complex seals may require multiple castings or additional components
\end{itemize}

\textbf{Duration:} Immediate effect with potential for temporary bypass (1 day)

\textbf{Casting Time:} Some Time (10-20 minutes depending on complexity)

\section*{The Feast of Shadows}
\textit{A ritual to draw power from darkness and despair}

\textbf{Purpose:} Gain strength by feeding on negative emotions or dark environments

\textbf{Primary Caster:} Spirit 4+, Arcana 3+, Presence 2+

\textbf{Helpers:} 1-2 participants with Spirit or Intimidation skills

\textbf{Channel Phase (Spirit + Arcana):}
\begin{enumerate}
\item Position yourself in an area of darkness, fear, or despair
\item Prepare to channel the negative energy through ritual focus
\item Roll Spirit + Arcana for primary caster, Spirit + Intimidation for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Command):}
\begin{enumerate}
\item Draw the negative emotions or ambient darkness into yourself
\item Transform the darkness into usable magical energy
\item Roll Presence + Command for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The power drawn is tainted; -1 die to all social rolls for one session as negative emotions cling to the caster
\item \textbf{3-4 SB:} The ritual attracts entities that feed on the same darkness; 2 SB added to next supernatural encounter
\item \textbf{5+ SB:} The caster becomes addicted to the dark power; must make Spirit + Resolve (DV 3) rolls to resist using the ritual again within one week
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Gain temporary magical power based on the level of darkness/despair present
\item Restore 1 Fatigue per success level, up to maximum Fatigue capacity
\item Gain +1 die to all rolls for Significant Time (duration based on power drawn)
\item Can enhance one magical effect per level of power drawn
\item Power fades gradually over several hours unless sustained by continued exposure
\end{itemize}

\textbf{Duration:} Temporary boost lasting 1-3 hours based on power drawn

\textbf{Casting Time:} Some Time (requires immersion in dark environment)

\section*{The Crown of Seasons}
\textit{A ritual to attune oneself to natural cycles and weather patterns}

\textbf{Purpose:} Gain power and insight aligned with natural cycles and seasonal changes

\textbf{Primary Caster:} Spirit 4+, Lore 3+, Survival 3+

\textbf{Helpers:} 2-3 participants with Survival, Lore, or Nature skills

\textbf{Channel Phase (Spirit + Lore):}
\begin{enumerate}
\item Meditate outdoors and observe the current natural conditions
\item Gather seasonal materials (flowers, leaves, stones, water)
\item Roll Spirit + Lore for primary caster, Wits + Survival/Lore for helpers
\end{enumerate}

\textbf{Weave Phase (Presence + Survival):}
\begin{enumerate}
\item Weave the seasonal materials into a crown or ceremonial focus
\item Align your spirit with the current season's power
\item Roll Presence + Survival for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The attunement is imperfect; gain only half the normal benefits for this season
\item \textbf{3-4 SB:} The ritual attracts the attention of local nature spirits who demand tribute; 1 SB added to next encounter with natural creatures
\item \textbf{5+ SB:} The caster becomes too attuned to nature, losing touch with civilization; -1 die to all social rolls involving urban environments for one session
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Gain seasonal attunement for one week (effects vary by season)
\item Spring: +1 die to healing and growth-related magic, +1 to all recovery rolls
\item Summer: +1 die to fire and light-based magic, resistance to heat effects
\item Autumn: +1 die to divination and knowledge magic, enhanced perception
\item Winter: +1 die to ice and preservation magic, resistance to cold effects
\item Gain +1 die to Survival rolls while in natural environments
\end{itemize}

\textbf{Duration:} One week, renewable with daily 5-minute meditation

\textbf{Casting Time:} Significant Time (requires outdoor ceremony and material gathering)

\section*{The Vessel of Echoes}
\textit{A ritual to store and replay magical effects or memories}

\textbf{Purpose:} Capture a magical effect, memory, or experience for later use

\textbf{Primary Caster:} Wits 4+, Arcana 4+, Lore 2+

\textbf{Helpers:} 1 participant with Lore or Insight skills

\textbf{Channel Phase (Wits + Arcana):}
\begin{enumerate}
\item Prepare the storage vessel (crystal, vial, focus object)
\item Determine what specific effect, memory, or experience to capture
\item Roll Wits + Arcana for primary caster, Wits + Lore/Insight for helper
\end{enumerate}

\textbf{Weave Phase (Spirit + Lore):}
\begin{enumerate}
\item Channel the target into the prepared vessel through focused will
\item Seal the vessel with appropriate binding rituals
\item Roll Spirit + Lore for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The storage is unstable; the captured effect has a 50\% chance of failing when replayed
\item \textbf{3-4 SB:} The vessel leaks energy; 1 SB added to next magical encounter as residual magic attracts attention
\item \textbf{5+ SB:} The vessel becomes cursed, replaying the captured effect at random times; caster suffers unexpected activation once per day
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Store one magical effect, memory, or experience in a prepared vessel
\item Stored effect can be replayed once by anyone holding the vessel
\item Memories can be shared with others through the vessel (as if experiencing them firsthand)
\item Complex effects may require multiple vessels or increase backlash risk
\item Vessel remains active for one month or until used
\end{itemize}

\textbf{Duration:} One month or until activated

\textbf{Casting Time:} Some Time (15-30 minutes of careful preparation)

\section*{The Thread of Fates}
\textit{A ritual to glimpse the connections between people and events}

\textbf{Purpose:} Perceive the invisible threads that bind individuals to each other and to destiny

\textbf{Primary Caster:} Wits 5+, Arcana 4+, Insight 3+

\textbf{Helpers:} 1-2 participants with Insight or Investigation skills

\textbf{Channel Phase (Wits + Insight):}
\begin{enumerate}
\item Gather personal items or samples from all individuals to be examined
\item Prepare the divination space with thread, yarn, or similar materials
\item Roll Wits + Insight for primary caster, Wits + Insight/Investigation for helpers
\end{enumerate}

\textbf{Weave Phase (Spirit + Arcana):}
\begin{enumerate}
\item Cast the threads across the divination space, allowing them to settle naturally
\item Interpret the patterns formed by the threads' connections
\item Roll Spirit + Arcana for primary caster
\end{enumerate}

\textbf{Backlash:}
\begin{itemize}
\item \textbf{1-2 SB:} The connections shown are partially obscured or misleading; -1 die to subsequent Investigation rolls for one scene
\item \textbf{3-4 SB:} The caster glimpses connections they would rather not see; suffer Fatigue 1 and -1 die to social rolls for Significant Time
\item \textbf{5+ SB:} The ritual reveals a connection that puts the caster in danger; GM introduces a threat linked to one of the revealed connections
\end{itemize}

\textbf{Effect:}
\begin{itemize}
\item Visualize the major connections between 2-5 specified individuals
\item See 1-2 significant past events that connect these individuals
\item Gain insight into one potential future connection or meeting
\item Identify one hidden relationship or secret bond between the subjects \item Gain +2 dice to Investigation or Social rolls involving the connected individuals for this session
\end{itemize}

\textbf{Duration:} Vision lasts for this scene; knowledge benefits last for this session

\textbf{Casting Time:} Some Time (requires setup and meditation)

\section*{The Ending}
These rituals represent the known arts of the Tulkani tradition as recorded in the surviving fragments. Many more surely existed, lost to time, war, or deliberate destruction.

The path of magic is ever one of discovery and rediscovery. What was once known may be found again, and what was never known may yet be revealed to those brave enough to seek beyond the veil.

Use these rituals as a foundation for your own explorations. Each caster must find their own way through the intricate paths between the worlds, learning what works and what warns them away through experience.

Remember always that magic is not without cost. The greatest powers demand the greatest sacrifices, and even the smallest rituals leave their mark upon both caster and world.

May your paths be safe, your intentions clear, and your wisdom greater than your ambition.

% !TEX root = resource_guide_main.tex
% Resource Guide Subsection: Disabling Magical Traps (Wards)
\chapter{Disabling Traps and Wards}\label{chapter:disabling-wards}
\index{Wards}\index{Traps}\index{DV}\index{Position}\index{Clocks}\index{Story Beats}

Traps are mechanical devices designed to either act as a [BARRIER] or do damage (+Fatigue for poison, or +Harm).
Wards are traps encoded in magic: a binding, sensor, or anchored effect. Treat them like any obstacle: \textbf{state the fiction}, \textbf{set Position}, and \textbf{resolve against DV}. This subsection expands the SRD example with GM-facing tools: how to set DV from fiction, what good/bad results look like, and how to keep the scene moving.

\subsection{Structure (at the table).}
\begin{enumerate}[label=\textbf{Step~\arabic*:}, leftmargin=2.1em]
\item \textbf{Name the Ward.} What does it sense (heat, oath-breaking, Patron presence)? What does it do (alarm, bind, blast, invert gravity)? Where is the \emph{anchor} (sigil, keystone, idol, threshold)?
\item \textbf{Reveal Handles.} Offer at least two fictional levers: the arcane pattern (\emph{Arcana}), the mechanism/anchor (\emph{Tinker}), and the approach route (\emph{Agility/Move}). Tools, components, or Patron symbols count as handles.
\item \textbf{Set Position \textrightarrow{} DV.} Use the ladder as defaults; shift for scale/urgency.
\begin{itemize}
\item \textbf{Controlled} (time, tools, clear access): \textbf{DV2}.
\item \textbf{Risky} (pressure, partial access): \textbf{DV3}.
\item \textbf{Desperate} (activating, hostile field): \textbf{DV~4–5+}.
\end{itemize}
\item \textbf{Declare Stakes.} What happens on Partial/Miss? Name the \emph{Ward Clock} and the \emph{Alert Clock} up front if relevant.
\end{enumerate}

\subsection{Common Ward Types & Anchors.}
\begin{tabularx}{\linewidth}{>{\bfseries}l >{\raggedright\arraybackslash}X >{\raggedright\arraybackslash}X}
\toprule
Type & Trigger & Anchor Examples \\
\midrule
Sentry & Crossing a threshold, line of sight, sound cadence & Lintel sigil, eye-gem, bell filament \\
Binding & Handling an item, touching a seal, oath breach & Wax seal, chain knot, iron brand \\
Punitive & Trespass + time window, named rival, taboo action & Idol tooth, ash sigil under rug, curse knot \\
Displacement & Path inversion, gravity twist, dream-step & Floor mandala, mirrored post, sleep pin \\
Anti-Magic & Casting in zone, symbol presence, boon use & Null plate, cold-iron frame, void ampoule \\
\bottomrule
\end{tabularx}

\subsection{Approaches & Aids (examples).}
\begin{itemize}
\item \textbf{Wits + Arcana:} analyze glyph grammar; ground the loop with salt/ink; counter-frequency hum.
\item \textbf{Wits + Tinker:} wedge the keystone; snip hair-thin filament; re-route power to a safe sink.
\item \textbf{Body + Agility:} thread the safe arc; pin the pressure plate; time the chime cadence.
\item \textbf{Teamwork:} one character steadies Position with tools while another rolls; assistants may accept \textbf{1~SB} to grant +1d once.
\item \textbf{Talents/Tools:} lockpicks, ward-ink, tuning forks, mirror-shards; Patron symbols might improve Position if thematically aligned.
\end{itemize}

\subsection{Clocks that Matter.}
\begin{description}[leftmargin=1.6em]
\item[\textsc{Ward Integrity} (4):] how close the binding is to disarming \emph{or} rupturing. On Partials, tick 1; on Misses, tick 2.
\item[\textsc{Alerted Attention} (4):] nearby guardians, spirits, or owners notice tampering.
\item[\textsc{Escalation} (4/6):] punitive effects ramp (more Harm, broader area, additional tags).
\end{description}

\subsection{Outcomes Palette (use or adapt).}
\begin{itemize}
\item \textbf{Strong Hit:} Ward disabled/redirected cleanly; optionally \emph{bank a tell} (learn its signature for later).
\item \textbf{Mixed:} Suppressed but \emph{unstable}; pick one: tick \textsc{Alerted Attention}~+1, apply minor Backlash by element/Patron, or leave a trace (future complication).
\item \textbf{Miss:} Ward holds or triggers; GM may (a) advance \textsc{Ward Integrity} toward rupture, (b) trigger a reduced effect now, (c) spend SB to introduce a hard complication.
\end{itemize}

\paragraph{Counter-Ward (Arcana Focus; optional).}
On a Success, choose one: (1) \textbf{Suppress} for a scene; (2) \textbf{Disarm} permanently (mark +1 Obligation or expend rare material); (3) \textbf{Invert} (redirect first trigger to a designated sink; start \textsc{Unintended Echo}~1/4).

\subsection{Difficulty Dials.}
\begin{itemize}
\item \textbf{Scale:} Large areas or ancient wards add +1 DV or an extra clock.
\item \textbf{Haste:} Rushed attempts worsen Position one step.
\item \textbf{Right of Names:} Knowing the warder’s true name improves Position one step or grants +1d.
\item \textbf{Wrong Patron:} Symbols from an opposed Patron impose –1d or +1 DV (GM call).
\end{itemize}

\subsection{Fiction First Hooks.}
\begin{itemize}
\item Show sensory tells (ozone, cold breath, hair standing) before rolls.
\item On Partials, change the room: new cover lines, heat shimmer, dead zones for boons.
\item On Miss, give a choice: suffer a lesser effect now to avoid full trigger later, or press on and risk rupture.
\end{itemize}

\paragraph{Worked Mini-Example.}
\emph{Vault Sentry (DV 3, Risky).} Anchor: eye-gem in an idol. The crew uses Wits+Tinker with ward-ink and a mirror shard. Mixed result: \textsc{Ward Integrity}+1, \textsc{Alerted Attention}+1, but the beam is shunted for this scene. Later a rival recognizes the mirror signature—payoff for the trace left behind.

\clearpage

\chapter{Poison as Story Beats (Narrative \texorpdfstring{$\leftrightarrow$}{↔} Mechanics)}\label{sec:poison-sb}
\index{Poison}\index{Story Beats}\index{Fatigue}\index{Harm}\index{Clocks}\index{Conditions}

Poison is best when it \emph{pushes scenes forward} rather than only dealing damage. This module frames poison as \textbf{Fatigue-forward}, with \textbf{Story Beats (SB)} powering escalation.

\subsection*{Core Model}
\begin{itemize}
\item \textbf{Primary Effect:} Poison applies \textbf{Fatigue} (not Harm) unless it explicitly says otherwise.
\item \textbf{Escalation:} The GM may \textbf{spend 1 SB} created in the scene to upgrade a poison instance: \emph{Fatigue} (\to) \emph{Condition} (\to) \emph{Harm}. Name the fiction of the escalation (spasms, fever spike, airway closing).
\item \textbf{Clocks:} Use a named clock per poison (e.g., \textsc{Neurotoxin Load} 1/4). Ticks convert into Fatigue/Conditions at thresholds.
\item \textbf{Resist:} Players may Resist poison normally; strong fiction (antidote, mask, purge rite) can reduce Position or cancel a tick.
\end{itemize}

\subsection*{Delivery \& Position}
\begin{tabularx}{\linewidth}{>{\bfseries}l l X}
\toprule
Route & Base DV & Position Cues \\
\midrule
Contact & 2 & Sweat, numbness, rashes; Controlled with gloves/time. \\
Ingested & 3 & Bitter note, wine film; Risky at table, Controlled in lab. \\
Inhaled & 3--4 & Perfume cloud, dust motes; Desperate in confined spaces. \\
Injury/Venom & 3--4 & Fanged strike, tainted blade; Risky unless braced/prepared. \\
\bottomrule
\end{tabularx}

\subsection*{Outcomes Palette}
\begin{description}[leftmargin=1.6em]
\item[\textbf{Strong Hit.}] Avoid or purge: clear \textsc{Poison Clock} by 1 and ignore the current instance.
\item[\textbf{Mixed.}] Mark \textbf{Fatigue +1} \emph{or} take a minor \textbf{Condition} (\textsc{Dizzy}, \textsc{Shaking}). GM may tick the \textsc{Poison Clock} +1.
\item[\textbf{Miss.}] Mark \textbf{Fatigue +1} and tick \textsc{Poison Clock} +1–2. The GM may spend SB to apply a stronger Condition \emph{or} escalate to \textbf{Harm 1}.
\end{description}

\subsection*{Conversion \& Interactions}
\begin{itemize}
\item \textbf{Fatigue → Harm.} Normal Fatigue overflow still converts per your health rules (see \S\ref{sec:health-fatigue-harm-rev}).
\item \textbf{Armor.} Armor rarely helps vs. contact/ingested poisons; vs. \emph{injury} delivery, apply armor before poison unless the blade is \textbf{Barbed/Venomous} (tag), which bypasses one step of conversion.
\item \textbf{Rites.} \emph{Purge} (\S\ref{sec:universal-rituals}) clears 1–2 ticks or downgrades Condition \textrightarrow{} Fatigue (GM sets Obligation).
\item \textbf{Patrons.} Poison from a Patron-aligned source may trigger thematic backlash or Claims when resisted.
\end{itemize}

\subsection*{SB Dials for GMs}
\begin{itemize}
\item \textbf{Pressure Valve (soft).} Spend SB to \emph{delay} catastrophic onset: tick clock instead of immediate Harm, but widen the threat.
\item \textbf{Spike (hard).} Spend SB to \emph{frontload} impact: apply a named Condition now (\textsc{Blind}, \textsc{Cramping}) or force a risky reposition.
\item \textbf{Echo.} Bank 1 SB to trigger a delayed complication (e.g., relapse during negotiations).
\end{itemize}

\subsection*{Sample Poisons}
\begin{tabularx}{\linewidth}{>{\bfseries}l X X}
\toprule
Name & Primary Effect (on hit) & SB-Driven Escalation \\
\midrule
Bitterglass & Fatigue +1; breath tastes like copper. & 1 SB: \textsc{Dizzy}. 2 SB: Harm 1 (air hunger). \\
Widowroot & Fatigue +1; numb grip. & 1 SB: drop held item. 2 SB: \textsc{Shaking}. \\
Dream-Marrow & Fatigue +1; slow recall. & 1 SB: \textsc{Fogged} (–1 to memory). 2 SB: \textsc{Asleep} unless shaken. \\
Serpent’s Borrow & Fatigue +1; swelling puncture. & 1 SB: \textsc{Cramping}. 2 SB: Harm 1 unless pressure bandaged. \\
Courtier’s Kiss & Fatigue +1; sweet aftertaste. & 1 SB: \textsc{Swayed} (–1 vs. social manipulation). 2 SB: \textsc{Blurred Vision}. \\
\bottomrule
\end{tabularx}

\subsection*{Worked Micro-Example}
\emph{Gala Balcony.} Inhaled perfume trap (DV~3, Risky). Mixed: PC marks Fatigue +1; GM ticks \textsc{Poison Clock}+1. Later, during the toast, the GM spends 1 SB to apply \textsc{Dizzy}; the PC chooses to lean on a friend (creating a bond) rather than push through and risk Harm.

\begin{tcolorbox}[title={Summary},colback=gray!5,colframe=black]
\small Poison leans on \textbf{Fatigue} first. \textbf{SB} lets the GM pace escalation as the scene demands. Name the clock, show the tells, and let fiction lead the effect.
\end{tcolorbox}

