% !TEX program = pdflatex
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=black}

\setlist{noitemsep,topsep=3pt,itemsep=2pt,parsep=0pt}
\titleformat{\section}{\large\bfseries}{}{0pt}{}

\begin{document}

% -------------------------------------------------------------
% HEADER
% -------------------------------------------------------------
\begin{center}
{\LARGE \textbf{Riarhin — Sapphire Wolf of the Northern Rims}}\\[4pt]
{\large Scout–Sharpshooter, Trusted Blade, Harbinger of Quiet Ends}
\end{center}

% -------------------------------------------------------------
% PROFILE
% -------------------------------------------------------------
\section*{Profile}
\begin{tabular}{@{}ll}
\textbf{Name:} & Riarhin \\
\textbf{Concept:} & Silent huntress and precision killer; bow of the Sapphire Wolf \\
\textbf{Tier:} & V (Legendary) \\
\textbf{Affiliation:} & Sapphire Wolf Company (Rim mercenaries), Ykrul tribunals \\
\textbf{Familiar:} & \textit{Ashwing} — Raven spirit bound by oath-sign, not Codex
\end{tabular}

\smallskip
\noindent
\emph{Riarhin holds no Patron.}  
Ikasha has laid a \emph{claim}, but she has not bent the knee, and no Codex binds her.  
Her gifts come from training, instinct, and something older than oaths—something the Sapphire Wolves whisper of in winter.

% -------------------------------------------------------------
% ATTRIBUTES
% -------------------------------------------------------------
\section*{Attributes}
\begin{tabular}{@{}ll}
\textbf{Body}  & 4 — impossible acrobatics, feral speed \\
\textbf{Wits}  & 3 — reading the field at a glance \\
\textbf{Spirit} & 3 — predator’s focus, unbreakable calm \\
\textbf{Presence} & 2 — she speaks with edges, not words
\end{tabular}

% -------------------------------------------------------------
% SKILLS
% -------------------------------------------------------------
\section*{Skills}
Archery 5; Stealth 4; Athletics 3; Awareness 3; Melee 3; Intimidation 2; Survival 3

% -------------------------------------------------------------
% TALENTS
% -------------------------------------------------------------
\section*{Talents}
\begin{itemize}
  \item \textbf{Custom Bow — \emph{Irisil}:} Ykrul-forged composite bow. When she has line-of-sight and patience, she acts with Dominant Position.
  \item \textbf{Deadeye:} On any attack from Far range, +1 Effect or knock target Prone on a Success.
  \item \textbf{Phantom Step:} Move from cover to cover without provoking Opportunity attacks on a Success.
  \item \textbf{Twin Dirks:} \emph{Toothpick} (Fate’s Edge stiletto) and \emph{The Last Blade} (Ykrul ritual blade). Gain +1 die when striking a surprised or shaken foe.
  \item \textbf{Sword — \emph{Remorse}:} When she draws Remorse, allies act with resolve; enemies must test Fear (DV set by GM). True power unknown.
  \item \textbf{Precision Kill:} On a roll of 10 while attacking from Far or High Ground, ignore armor for that strike.
  \item \textbf{Familiar — Ashwing: Beast Speech (1/scene):} Ashwing can speak briefly in mortal tongue; the raven delivers warnings, threats, commands, or eerie omens.
\end{itemize}

% -------------------------------------------------------------
% SIGNATURE WEAPONS
% -------------------------------------------------------------
\section*{Signature Kit}
\begin{itemize}
  \item \textbf{\emph{Irisil}, Sapphire Wolf Bow:} Reinforced horn and sinew, cold-tuned string. When setting an ambush, her first shot each scene counts as a Critical if it deals Harm.
  \item \textbf{\emph{Toothpick}:} Fine, razor stiletto; vanishes into sleeves and seams. Treat as Concealed until drawn.
  \item \textbf{\emph{The Last Blade}:} Ykrul Dirk; carried only for worthy kills. When it tastes blood, Stealth rolls gain +1 die for the next scene.
  \item \textbf{\emph{Remorse}:} Curved sword of silvered steel. Carried wrapped. Drawn only when judgment falls. When unsheathed, she may \emph{Silence} a battlefield: shouting becomes whispers; fear becomes stillness.
\end{itemize}

% -------------------------------------------------------------
% FAMILIAR
% -------------------------------------------------------------
\section*{Familiar — Ashwing, Raven Spirit}
Ashwing is bound by sign of the wolf-claw moon.  
Not a Patron’s emissary—something more primal, older than the Rims’ snow.

\textbf{Abilities:}
\begin{itemize}
  \item Scout from above; mark targets with cry or silence.
  \item Speak 1/scene in mortal tongue — clipped, cryptic, unmistakable.
  \item Sense killing intent; warn Riarhin with a rattle of wings.
  \item When ash falls or snow drifts, Ashwing can vanish into feathers of smoke.
\end{itemize}

\emph{Personality:} Mocking, hungry, fiercely protective.  
\emph{Bond:} ``Claw, bow, and blade. I choose the living storm.’’

% -------------------------------------------------------------
% COMBAT STYLE
% -------------------------------------------------------------
\section*{Combat Style — Wolf and Storm}
\begin{itemize}
  \item Riarhin never trades blows. She ends fights before they begin.
  \item When she acts from Far, she attacks twice: one real, one phantom—opponent must choose which to defend.
  \item If forced into melee, she moves like a serrated whisper. Dirks first; Remorse last.
\end{itemize}

\textbf{Wolf Sign:} When Sapphire Wolves fight at her side, all allies gain +1 die on the first violent action of a scene.

% -------------------------------------------------------------
% SPECIAL MOVES
% -------------------------------------------------------------
\section*{Special Moves (Tier V)}
\begin{itemize}
  \item \textbf{Silencer’s Blessing:} When she kills a target silently, she may erase their presence. Footprints fade. Blood cools to dust. No alarm.
  \item \textbf{High-String Shot:} Fire a heavy arrow into armor plates, locking joints or knocking shields aside.
  \item \textbf{Break the Line:} On a Critical from Far (10 on d10), she can nail two foes into cover with a single skewering shot.
  \item \textbf{Cold Oath:} If outnumbered, she gains +1 die to every attack until only one enemy remains.
\end{itemize}

% -------------------------------------------------------------
% RELATIONSHIPS
% -------------------------------------------------------------
\section*{Whispers of Ikasha}
Ikasha does not own her.  
But the runes of claim coil faintly along Remorse, unseen until the blade drinks the blood of tyrants.  
What that means… even Riarhin refuses to ask.

% -------------------------------------------------------------
% CONDITION & HARM
% -------------------------------------------------------------
\section*{Harm \& Resolve}
\begin{itemize}
  \item \textbf{Cold Blooded:} Ignore the first Level 1 Harm each scene.
  \item \textbf{When Level 2 Harm is taken:} She vanishes into motion; next attack gains +2 dice.
  \item \textbf{Level 3 Harm:} Unconscious but fingers still on Irisil. If Ashwing calls her name, she rises for one more shot.
\end{itemize}

% -------------------------------------------------------------
% RECOVERY SCENE
% -------------------------------------------------------------
\section*{Recovery Scene — Snow and Steel}
When resting, she sits in silence, sharpening Remorse or fletching arrows of blackened ash.  
If another warrior earns her respect, she teaches one silent kill or ghost-walk trick: that ally gains +1 die to Stealth next session.

% -------------------------------------------------------------
% END
% -------------------------------------------------------------
\end{document}