\chapter{Appendix: Tools and Tables}\index{appendix}

This appendix provides quick-reference tools, sample content, and interpretive guides to support your game. Whether you're running a quick session or a long campaign, these tables and tips will help you keep the world alive and the tension high.

\section*{Quick Reference Sheets}\index{quick reference}

\subsection*{Outcome Matrix}\index{Outcome Matrix}

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Case} & \textbf{Name} & \textbf{Guidance} \\
\midrule
$S \geq DV$ and $C = 0$ & Clean Success\index{Clean Success} & Deliver the intent crisply. \\
$S \geq DV$ and $C > 0$ & Success \& Cost\index{Success \& Cost} & Grant the intent; spend/bank CP for complications. \\
$0 < S < DV$ & Partial\index{Partial} & Progress with a fork. \\
	$S = 0$ & Miss\index{Miss} & No progress. Cash/bank CP. Award a Boon. \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Complication Point (CP) Spend Menu}\index{CP spend menu}

\begin{itemize}
    \item \textbf{1 CP}: Minor pressure: noise, trace, +1 Supply segment\index{Supply segment}.
    \item \textbf{2 CP}: Moderate setback: alarm raised, lose position/cover, lesser foe or lock.
    \item \textbf{3 CP}: Serious trouble: reinforcements, key gear breaks, rail tick\index{rail tick}.
    \item \textbf{4+ CP}: Major turn: trap springs, authority arrives, scene shifts.
\end{itemize}

\begin{table}[htbp]
\centering
\caption{Session Type Adaptation}
\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|p{4cm}|}
\hline
\textbf{Session Focus} & \textbf{Recommended Ratio} & \textbf{Pacing Priority} & \textbf{Application Notes} \\
\hline
Action-Oriented & 80/20 & Speed & Maintain combat flow; save deck draws for major turning points \\
\hline
Narrative-Focused & 60/40 & Depth & Emphasize thematic complications; rich story development \\
\hline
Investigation & 85/15 & Clarity & Quick resolution supports deduction; avoid overcomplication \\
\hline
Climactic & 50/50 & Impact & Maximum complication weight; story-defining moments \\
\hline
\end{tabular}
\end{table}

\subsection*{Position Descriptions}\index{position}

\begin{itemize}
    \item \textbf{Controlled}\index{position!Controlled}: You act on your terms.
    \item \textbf{Risky}\index{position!Risky}: You act under pressure.
    \item \textbf{Desperate}\index{position!Desperate}: The odds are stacked against you.
\end{itemize}

\subsection*{Difficulty Ladder (Set Before the Roll)}\index{Difficulty Ladder}

\begin{center}
\begin{tabular}{cll}
\toprule
\textbf{DV} & \textbf{Name} & \textbf{When to Use} \\
\midrule
1 & Routine\index{Difficulty Value!Routine} & Clear intent, modest stakes, controlled environment. \\
2 & Pressured\index{Difficulty Value!Pressured} & Time pressure, mild resistance, partial info. \\
3 & Hard\index{Difficulty Value!Hard} & Hostile conditions, active opposition, precise timing. \\
4+ & Extreme\index{Difficulty Value!Extreme} & Multiple constraints, high precision, dramatic failure. \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Sample NPCs}\index{NPCs}

\subsubsection*{Encounters}\index{NPCs!encounters}

\begin{itemize}
    \item \textbf{Bandit Skirmisher}: Body 2, Wits 2. Melee 2, Stealth 1. Light armor, opportunistic.
    \item \textbf{Ykrul Rider}: Body 4, Wits 3. Riding 3, Melee 3. Mobile, brutal.
    \item \textbf{Street Bravo}: Presence 3, Body 2. Dueling 3. Quick to anger.
\end{itemize}

\subsubsection*{Foils \& Rivals}\index{NPCs!foils}

\begin{itemize}
    \item \textbf{Ambitious Scribe}: Wits 3, Presence 3. Intrigue 3, Lore 2. Always knows a rumor.
    \item \textbf{Mercenary Captain}: Body 4, Spirit 3. Command 3, Melee 4. Pragmatic, dangerous ally.
    \item \textbf{Flame Preacher}: Presence 4, Spirit 3. Oratory 4, Faith 3. Incites mobs.
\end{itemize}

\subsubsection*{Prestige NPCs}\index{NPCs!prestige}

\begin{itemize}
    \item \textbf{High Elf Loremaster}: Wits 5, Spirit 4. Lore 5, Arcana 4. Knows secrets older than nations.
    \item \textbf{Dwarven Forge-Patriarch}: Body 5, Spirit 4. Craft 5, Command 4. Commands stone and steel.
    \item \textbf{Ykrul Warglord}: Body 5, Presence 4. Command 4, Melee 5. Unites clans with blood and will.
\end{itemize}

\section*{Deck of Consequences Interpretation Guide}\index{Deck of Consequences}

\subsection*{Two Deck Systems}\index{Deck of Consequences!two deck systems}

\paragraph{Deck of Consequences (scene drama).}
\emph{Hearts}\index{Hearts (suit)}=social fallout, \emph{Spades}\index{Spades (suit)}=harm/escalation, \emph{Clubs}\index{Clubs (suit)}=material cost, \emph{Diamonds}\index{Diamonds (suit)}=magical/spiritual disturbance.

\paragraph{Travel Decks (regional, 52-card).}\index{Travel Decks}
\emph{Spade}\index{Spades (suit)}=Place, \emph{Heart}\index{Hearts (suit)}=Actor, \emph{Club}\index{Clubs (suit)}=Pressure, \emph{Diamond}\index{Diamonds (suit)}=Leverage.

\subsection*{Hearts (Emotional/Social)}\index{Hearts (suit)}

\begin{itemize}
    \item \textbf{Ace--3}: Minor offense, awkward moment.
    \item \textbf{4--6}: Relationship strain, public embarrassment.
    \item \textbf{7--9}: Betrayal, scandal, loss of trust.
    \item \textbf{10--King}: Heartbreak, exile, shattered alliance.
\end{itemize}

\subsection*{Spades (Harm/Escalation)}\index{Spades (suit)}

\begin{itemize}
    \item \textbf{Ace--3}: Bruise, scrape, fatigue.
    \item \textbf{4--6}: Wound, gear damaged, position lost.
    \item \textbf{7--9}: Severe injury, ally down, structural collapse.
    \item \textbf{10--King}: Death, dismemberment, permanent loss.
\end{itemize}

\subsection*{Clubs (Material/Cost)}\index{Clubs (suit)}

\begin{itemize}
    \item \textbf{Ace--3}: Minor loss, delayed payment.
    \item \textbf{4--6}: Gear failure, debt incurred.
    \item \textbf{7--9}: Major asset lost, bankruptcy.
    \item \textbf{10--King}: Total ruin, legacy debt.
\end{itemize}

\subsection*{Diamonds (Magical/Spiritual)}\index{Diamonds (suit)}

\begin{itemize}
    \item \textbf{Ace--3}: Omen, whisper, strange coincidence.
    \item \textbf{4--6}: Curse triggered, spirit appears, past returns.
    \item \textbf{7--9}: Arcane backlash, forbidden knowledge revealed.
    \item \textbf{10--King}: Reality bends, godlike force intervenes.
\end{itemize}

\section*{Campaign Clock Examples}\index{Campaign Clocks}

\subsection*{Mandate Advancement Triggers}\index{Campaign Clocks!Mandate}

\begin{itemize}
    \item Public victory in battle or debate.
    \item Successful resolution of a major crisis.
    \item Recognition by a powerful faction or ruler.
\end{itemize}

\subsection*{Crisis Advancement Triggers}\index{Campaign Clocks!Crisis}

\begin{itemize}
    \item Rival faction gains influence or territory.
    \item Asset neglect or betrayal.
    \item Scandal or public loss of trust.
\end{itemize}

\section*{Travel Clock Sizes}\index{clocks!Travel}

\begin{itemize}
    \item \textbf{2--5}: 4 segments (short leg, low risk).
    \item \textbf{6--10}: 6 segments (standard journey).
    \item \textbf{J/Q/K}: 8 segments (long or dangerous route).
    \item \textbf{Ace}: 10 segments (epic or supernatural travel).
\end{itemize}

\section*{Follower and Asset Condition States}\index{Followers!condition}\index{Assets!condition}

\begin{itemize}
    \item \textbf{Maintained}: Full capability.
    \item \textbf{Neglected}: -1 die penalty; narrative wear.
    \item \textbf{Compromised}: Unavailable until repaired or recovered.
\end{itemize}

\section*{Boon Economy Quick Guide}\index{Boon economy}

\begin{itemize}
    \item \textbf{Holding cap}: You can hold at most 5 Boons\index{Boons}.
    \item \textbf{Conversion}: Once per session, in downtime, you may convert 2 Boons → 1 XP (max 2 XP via conversion per session).
    \item \textbf{Using Boons}: Re-roll one die after seeing the pool; Activate an Off-Screen Asset.
\end{itemize}

\section*{Mechanical Constraints}\index{mechanical constraints}

\begin{itemize}
    \item \textbf{ASSIST MAX}: +3 dice total per roll, regardless of helpers. Exception: The "Exceptional Coordination" Talent allows one follower to provide +4 assist dice.
    \item \textbf{BOON MAX}: 5 total, 2→1 XP conversion once/session (max 2 XP via conversion per session).
    \item \textbf{INITIATIVE}\index{Initiative Actions}: 1 Follower Action per scene party-wide.
    \item \textbf{OVER-STACK}\index{Over-Stack}: 2+ structural advantages = start rails +1 OR GM banks +1 CP.
    \item \textbf{POSITION}: Controlled | Risky | Desperate (affects success/failure texture).
\end{itemize}

\section*{Let the World Sing}\index{narrative first}

These tools are not meant to constrain your game---they are meant to \textbf{support your vision}. Use them to keep tension high, consequences real, and the story moving forward.

Because in the end, it is not the GM who writes the legend.

It is the players.

You simply hold the quill.

\end{chapter}
