% !TEX root = ../fates_edge_players_guide.tex

\chapter{Assets and Followers}
\label{ch:assets-followers}

In \textbf{Fate's Edge}, your influence extends beyond your own skills and talents. Through \textbf{Assets} \index{Assets} and \textbf{Followers} \index{Followers}, you can shape the world off-screen, command allies in battle, and build lasting legacies. But with power comes responsibility—and risk.

\section{Assets}
\index{Assets}

\textbf{Assets} are off-screen resources that solve problems between sessions or provide leverage during scenes. They do not act in scenes directly, but they change the fiction and provide mechanical or narrative benefits.

\subsection*{Types of Assets}
\index{Assets!types}

\begin{description}
  \item[Minor (4 XP, 1 day)] \index{Assets!Minor} — Safehouse, small shop, petty title, local contact network.
  \item[Standard (8 XP, 1 week)] \index{Assets!Standard} — Noble title, guild membership, spy ring, workshop, small vessel.
  \item[Major (12 XP, 1 month)] \index{Assets!Major} — City charter, regional network, fortress, major magical item, significant artifact.
\end{description}

\subsection*{Using Assets}
\index{Assets!usage}

Assets provide benefits:

\begin{itemize}
  \item \textbf{Off-Screen Effect:} Use each Asset's listed Off-Screen effect once per session for free.
  \item \textbf{On-Screen Activation:} To reshape the current scene, spend 1 Boon \index{Boons}. The Asset must have scope and reach for the intended effect.
  \item \textbf{Downtime Activation:} A player may activate an off-screen asset at the very start of a campaign or during Downtime. It costs 2 XP or 1 Boon to activate.
\end{itemize}

\subsection*{Asset Condition}
\index{Assets!condition}

All Assets have a \textbf{Condition Track} which reflects their operational status:

\begin{description}
  \item[Maintained] \index{Assets!condition!Maintained} — Full capability. Functions normally.
  \item[Neglected] \index{Assets!condition!Neglected} — Impaired. Impose a -1 die penalty when used; requires attention or XP to restore.
  \item[Compromised] \index{Assets!condition!Compromised} — Unavailable. Cannot be used until repaired or recovered (requires Downtime action or XP spend).
\end{description}

\section{Followers}
\index{Followers}

\textbf{Followers} are on-screen allies who assist in scenes. They are not stat blocks—they are story agents with personalities, goals, and flaws, represented by simple mechanics.

\subsection*{Follower Ratings}
\index{Followers!ratings}

Followers are rated by \textbf{Cap} (1–5), which determines their specialty dice pool and assist bonus.

\textbf{Cost:} A follower with Specialty Cap C costs C² XP. Downtime = 1–3 days to recruit and brief.

\subsection*{Assisting in Scenes}
\index{Followers!assistance}

Followers assist by adding dice to your rolls:

\begin{itemize}
  \item Assist dice come from the helper, not the leader.
  \item Total Assist on any roll (from any sources) remains hard-capped at +3. Exception: The "Exceptional Coordination" \index{Exceptional Coordination} Talent allows one follower to provide +4 assist dice.
  \item When applicable, the follower adds help dice equal to \textbf{min(Cap, the helper's relevant Skill)}, capped at +3 dice.
  \item \textbf{Slot Limit:} Only one follower may assist a given action.
\end{itemize}

\subsection*{Follower Condition}
\index{Followers!condition}

Followers track \textbf{Exposure} and \textbf{Harm} to represent their well-being:

\begin{description}
  \item[Exposure] \index{Followers!Exposure} — Heat, attention, stress, overextension, or narrative pressure placed upon the follower.
  \item[Harm] \index{Followers!Harm} — Physical injury, trauma, fatigue, or direct damage to the follower.
\end{description}

\textbf{States:}
\begin{itemize}
  \item \textbf{Maintained} \index{Followers!condition!Maintained} — Reliable and ready.
  \item \textbf{Neglected} \index{Followers!condition!Neglected} — Needs downtime or care. Impose a -1 die penalty to their assistance.
  \item \textbf{Compromised} \index{Followers!condition!Compromised} — Captured, defected, lost, or incapacitated. Cannot assist until recovered (requires Downtime action or specific fiction).
\end{itemize}

\textbf{Tracks:} Exposure and Harm typically use a simple track: 0 (None) $\rightarrow$ 1 (Light) $\rightarrow$ 2 (Moderate) $\rightarrow$ 3 (Severe) $\rightarrow$ 4 (Critical). At Critical, the follower is typically Out of Action or Lost.

\subsection*{Upkeep}
\index{Followers!upkeep}

Maintaining followers requires effort:

\begin{itemize}
  \item \textbf{Downtime Cost:} Each Downtime, pay XP equal to the follower's Cap or spend a Scene tending the relationship (roleplaying interaction) to maintain their Condition.
  \item \textbf{Risk of Harm/Exposure:} If the GM spends 2+ Story Beats \index{Story Beat (SB)} on an action you take with a follower's assistance, they may mark Exposure or Harm on the follower instead of applying other consequences, if fictionally appropriate.
  \item \textbf{Off-Screen Capability:} Once per downtime, a follower with Cap 3 or higher can attempt to solve one significant off-screen problem but generates 1 SB for the party. The GM must describe how the follower's action creates story consequences for the SB generated.
\end{itemize}

\section{Follower Initiative Actions}
\index{Followers!initiative actions}

Once per scene (across the entire party), one on-screen follower may take a small independent action:

\begin{itemize}
  \item \textbf{Scout \& Signal} — Change an ally's next action position to Controlled.
  \item \textbf{Distract \& Draw} — Reduce a kinetic rail (Hunt/Escape/Hazard) by –1 tick.
  \item \textbf{Fetch \& Carry} — Move a small object through danger.
\end{itemize}

\textbf{Cost:} To take an Initiative Action, mark Exposure +1 or Harm 1 on that follower.

\section{Loyalty \& Bonds (Optional Rules)}
\index{Followers!loyalty}
\index{Followers!bonds}

\begin{itemize}
  \item \textbf{Loyalty Tag:} Track a simple Loyalty tag per follower: Wary / Steady / Devoted.
  \item \textbf{Devoted:} Devoted followers can once per arc convert one GM Complication targeting them into a lesser setback (e.g., a 3 SB spend becomes a 1 SB spend).
  \item \textbf{Wary:} Wary followers cost +1 XP to Maintain and may require a successful roll (e.g., Presence + Sway, DV 2) to assist effectively.
  \item \textbf{Bonds:} Followers can be tied to character Bonds. A follower with a Bond to a PC might gain a +1 die bonus to assist that PC, or the PC might gain a Boon by aiding the follower in a significant action.
\end{itemize}

\section{Follower Examples}
\index{Followers!examples}

\textbf{Cap 1 Follower:} \textit{Street Urchin} \\
Skills: Move 2, Notice 1 \\
Assist: Provides +1 die to Stealth or Investigation in urban environments. \\
Quirk: Easily startled, might flee if exposed to violence.

\textbf{Cap 3 Follower:} \textit{Mercenary Sergeant} \\
Skills: Combat 3, Tactics 2, Physique 2 \\
Assist: Provides up to +3 dice to Combat or Tactics rolls. Can act as a secondary guard. \\
Quirk: Demands a share of loot and may question orders.

\textbf{Cap 5 Follower:} \textit{Ship Captain} \\
Skills: Pilot 4, Combat 3, Sway 3, Lore 2 \\
Assist: Provides up to +5 dice (capped at +3) to relevant rolls involving ships, navigation, or crew management. \\
Quirk: Has their own agenda related to trade routes or personal vendettas.

\section{Summary}

Assets and Followers are powerful tools—but they come with obligations:

\begin{itemize}
  \item Assets solve problems off-screen but require maintenance and can be Compromised.
  \item Followers assist in scenes but are vulnerable to Complications, Exposure, and Harm.
  \item Both shape the story and reflect your character's growing influence and responsibilities.
\end{itemize}

Invest wisely—and be ready to pay the price for the power you wield through others.

\end{chapter}
