% !TEX root = ../fates_edge_players_guide.tex

\chapter{Gods, Powers, and Patrons}
\label{ch:gods-powers-patrons}

In \textbf{Fate's Edge}, religion is not just faith—it is \textbf{faction}. The gods, powers, and ascended entities of the world are active forces, shaping cultures, inspiring zealots, and demanding tribute. This chapter explores the major divine and mystical forces, their followers, and how they influence the world. It also details the powerful, often alien Patrons who offer Rites in exchange for attention and obligation.

\section{Faith as Faction}
\index{Faith as faction}

Religion in Fate's Edge is deeply political. Each power has its own churches, cults, schisms, and agendas. To serve a god is to join a faction—with all the alliances, enemies, and obligations that come with it. Over time, many former deities have transcended their personal identities, becoming abstract forces aligned to their sphere of influence, known as \textbf{Patrons}.

\section{Major Powers}
\index{Powers!major}

\subsection*{Everflame}
\index{Powers!Everflame}

\textbf{Domains:} Fire, purification, law \\
\textbf{Seat:} Ecktoria \\
\textbf{Themes:} Order, sacrifice, judgment \\
\textbf{Patron Aspect:} The Oath of Flame and Light

The Everflame is the dominant faith of the northern shores. It teaches that truth burns away falsehood, and that only through suffering can the soul be purified. Its priests are inquisitors, judges, and executioners.

\textbf{Cleric Talents:} Everflame Keeper, Zealot's Blade, Purifying Light

\subsection*{Light}
\index{Powers!Light}

\textbf{Domains:} Revelation, mercy, community \\
\textbf{Seat:} Viterra \\
\textbf{Themes:} Healing, unity, evangelism \\
\textbf{Patron Aspect:} The Oath of Flame and Light

A reformation of the Everflame, the Light emphasizes compassion and community over punishment. Its followers are healers, teachers, and missionaries who seek to unite the faithful under a single banner.

\textbf{Cleric Talents:} Healing Light, Community Blessing, Evangelist's Tongue

\subsection*{Khemesh and Ráeyn}
\index{Powers!Khemesh and Ráeyn}

\textbf{Domains:} Krakens, storms, the sea \\
\textbf{Seat:} Kahfagia \\
\textbf{Themes:} Power, survival, dominion \\
\textbf{Patron Aspects:} Khemesh, The Kraken; Raéyn of the Sea

The twin patrons of Kahfagia, Khemesh and Ráeyn are worshipped by sailors, pirates, and storm-riders. Their temples are built on cliffs and islands, and their priests read the will of the deep in the bones of whales.

\textbf{Cleric Talents:} Storm's Favor, Kraken's Blessing, Deep Oracle

\subsection*{Dhahar, Lord of the Sky and War}
\index{Powers!Dhahar}

\textbf{Domains:} Sky, war, honor \\
\textbf{Seat:} Dhahara \\
\textbf{Themes:} Battle, banners, judgment \\
\textbf{Patron Aspect:} Dhahar (Note: Dhahar is a traditional deity, less commonly accessed via Rites)

Dhahar is the god of the open sky and righteous war. His followers swear by banners and settle disputes with steel. In the south, to break an oath sworn under his name is to invite his wrath.

\textbf{Cleric Talents:} Banner of War, Sky's Judgment, War-Chanter

\subsection*{Ikasha, She Who Sleeps Within Shadow}
\index{Powers!Ikasha}

\textbf{Domains:} Shadow, secrets, transformation \\
\textbf{Seat:} Tulkani underground \\
\textbf{Themes:} Mystery, change, forbidden knowledge \\
\textbf{Patron Aspect:} Ikasha, She Who Sleeps

Ikasha is whispered more than preached. Her followers are spies, assassins, and those who walk between worlds. Her temples are hidden, and her rites are known only to the initiated. She is one of the most prominent Patrons.

\textbf{Cleric Talents:} Shadow Walk, Secret Keeper, Veil of Transformation

\subsection*{The Old Ancestors (Dwarven Faith)}
\index{Powers!Old Ancestors}

\textbf{Domains:} Kin, stone, memory \\
\textbf{Seat:} Aeler mountain halls \\
\textbf{Themes:} Tradition, craft, lineage \\
\textbf{Patron Aspect:} The Old Ancestors (Note: Ancestral veneration is less structured as Patron Rites)

The dwarves venerate their ancestors as guardians of the deep laws. To speak an ancestor's name is to invoke their protection—and their expectations.

\textbf{Cleric Talents:} Ancestral Guidance, Stone Communion, Kin-Binding Oath

\section{Other Powers and Spirits}
\index{Powers!other}

\begin{description}
  \item[Ykrul Sky-Spirit] \index{Powers!Ykrul Sky-Spirit} — Nameless shamanic power tied to omens and war.
  \item[Valewood Entities] \index{Powers!Valewood Entities} — Ancient forest powers that shift with the seasons.
  \item[Fae Courts] \index{Powers!Fae Courts} — Elusive beings of whim and consequence.
  \item[The Hunger Below] \index{Powers!Hunger Below} — Primordial force stirring beneath the world.
\end{description}

\section{Patrons: Ascended Forces}
\index{Patrons}

Many former gods and powerful entities have transcended their personal personas, becoming abstract forces or offices aligned to a specific sphere. These \textbf{Patrons} are not merely worshipped; they are bargained with. They offer \textbf{Rites}—precise, named effects—in exchange for attention, marked as \textbf{Obligation}.

\subsection*{Nature of Patrons}
\begin{itemize}
    \item \textbf{Forces, Not Persons:} Patrons are aspects of reality, not personalities. They respond to correct procedures and fulfill their conceptual roles.
    \item \textbf{Rites over Prayer:} Interaction is through specific Rites, not general prayer or sacrifice.
    \item \textbf{Obligation:} Using a Rite increases your Obligation ledger with that Patron. High Obligation brings their attention, often as complications.
\end{itemize}

\subsection*{Examples of Patrons}
\begin{description}
    \item[The Witness] \index{Patrons!The Witness} — Observation, Truth, Surveillance.
    \item[The Sacred Geometry] \index{Patrons!The Sacred Geometry} — Order, Pattern, Structure.
    \item[The Sealed Gate] \index{Patrons!The Sealed Gate} — Boundaries, Jurisdiction, Closure.
    \item[The Traveler] \index{Patrons!The Traveler} — Ways Between, Roads, Journeys.
    \item[The Outsider] \index{Patrons!The Outsider} — Elsewhere, Intrusion, Planar Forces.
    \item[Mor’iraath, Demon of Fire] \index{Patrons!Mor’iraath} — Consumption, Passion, Destruction.
    \item[Mykkiel, Angel of Law] \index{Patrons!Mykkiel} — Edict, Process, Justice.
    \item[Isoka, Angel of Serpents] \index{Patrons!Isoka} — Change, Shedding, Transformation.
    \item[Inaea, Angel of the Spider] \index{Patrons!Inaea} — Webs, Fate, Connections.
    \item[Varnek Karn, the Bone King] \index{Patrons!Varnek Karn} — Death, Dominion, Undeath.
    \item[Mab, She of the Thorned Crown] \index{Patrons!Mab} — Twilight Bargains, Fae Pacts.
    \item[Nidhoggr, the World-Worm] \index{Patrons!Nidhoggr} — Dreaming Antiquity, Deep Time.
    \item[Maelstreus, the Red Prince] \index{Patrons!Maelstreus} — Infernal Deals, Temptation.
    \item[The Carrion-King] \index{Patrons!The Carrion-King} — Decay, Cycles, Rot.
    \item[The Gallow’s Bell] \index{Patrons!The Gallow's Bell} — Consequences, Retribution, Judgment.
    \item[The Clockwork Demon] \index{Patrons!The Clockwork Demon} — Forbidden Tech, Entropy.
\end{description}

\section{Divine Service and Consequences}
\index{Divine service}

Serving a god or engaging with a Patron in Fate's Edge is not without cost:

\begin{itemize}
  \item \textbf{Obligations} — Divine mandates, rituals, tithes, or Patron Rite ledgers.
  \item \textbf{Complications} — Gods may demand sacrifices or intervene in dramatic ways. Patrons manifest attention through complications when Obligation is high.
  \item \textbf{Blessings} — Divine favor grants Talents, Boons, or narrative edges. Patrons grant Rites.
\end{itemize}

\section{Cleric, Divine Spellcasting, and Rites}
\index{Spellcasting!divine}
\index{Rites}

Characters have multiple paths to interact with the divine and numinous:

\textbf{Faith-based Talents:} Traditional divine service grants access to Cleric Talents.

\textbf{Thaumaturgy:} A form of spellcasting tied to divine will rather than arcane study.
\begin{itemize}
  \item Purifying Light
  \item Divine Shield
  \item Blessing of Courage
  \item Wrath of the Heavens
\end{itemize}

\textbf{Pact Magic (Rites):} Bargaining with Patrons for specific, reliable effects.
\begin{itemize}
    \item Requires a \textbf{Thiasos Bond} (Familiar) and/or a \textbf{Codex of the Binding Art} (Asset).
    \item Invoked by spending \textbf{Boons}.
    \item Create \textbf{Obligation} which, if unchecked, leads to Patron attention.
\end{itemize}

\section{Summary}

The gods and forces of Fate's Edge, through their worshippers and pacted, are active players in the world:

\begin{itemize}
  \item Faith is faction, with politics and power struggles. Patrons are forces to be bargained with.
  \item Divine service and pacting bring both blessings and burdens.
  \item Gods and Patrons shape culture, law, and the fate of nations.
  \item Rites offer powerful, reliable effects at the cost of growing Obligation.
\end{itemize}

Choose your faith and your Patrons wisely—and be prepared to pay their price.

\end{chapter}
