% !TEX root = ../fates_edge_players_guide.tex

\chapter{Assets and Followers}
\label{ch:assets-followers}

In \textbf{Fate's Edge}, your influence extends beyond your own skills and talents. Through \textbf{Assets} \index{Assets} and \textbf{Followers} \index{Followers}, you can shape the world off-screen, command allies in battle, and build lasting legacies. But with power comes responsibility—and risk.

\section{Assets}
\index{Assets}

\textbf{Assets} are off-screen resources that solve problems between sessions. They do not act in scenes directly, but they change the fiction and provide leverage when you return to the table.

\subsection*{Types of Assets}

\begin{description}
  \item[Minor (4 XP, 1 day)] \index{Assets!Minor} — Safehouse, small shop, petty title.
  \item[Standard (8 XP, 1 week)] \index{Assets!Standard} — Noble title, guild section, spy ring.
  \item[Major (12 XP, 1 month)] \index{Assets!Major} — City license, regional network, fortress lease.
\end{description}

\subsection*{Using Assets}

Assets provide off-screen benefits:

\begin{itemize}
  \item Use each Asset's listed Off-Screen effect once per session for free.
  \item To reshape the current scene, spend 1 Boon \index{Boons}.
  \item The Asset must have scope and reach.
\end{itemize}

\subsection*{Asset Condition}
\index{Assets!condition}

All Assets have a \textbf{Condition Track}:

\begin{description}
  \item[Maintained] — Full capability.
  \item[Neglected] — -1 die when used; requires attention.
  \item[Compromised] — Unavailable until repaired or recovered.
\end{description}

\section{Followers}
\index{Followers}

\textbf{Followers} are on-screen allies who assist in scenes. They are not stat blocks—they are story agents with personalities, goals, and flaws.

\subsection*{Follower Ratings}
\index{Followers!ratings}

Followers are rated by \textbf{Cap} (1–5), which determines their specialty dice and assist bonus.

Cost: A follower with Specialty Cap C costs C² XP. Downtime = 1–3 days to recruit and brief.

\subsection*{Assisting in Scenes}
\index{Followers!assistance}

Followers assist by adding dice to your rolls:

\begin{itemize}
  \item Assist dice come from the helper, not the leader.
  \item Total Assist on any roll (from any sources) remains hard-capped at +3. Exception: The "Exceptional Coordination" \index{Exceptional Coordination} Talent allows one follower to provide +4 assist dice.
  \item When applicable, the follower adds help dice equal to \textbf{min(C, the helper's relevant Skill)}, capped at +3 dice.
  \item Slot Limit: Only one follower may assist a given action.
\end{itemize}

\subsection*{Follower Condition}
\index{Followers!condition}

Followers track \textbf{Exposure} and \textbf{Harm}:

\begin{description}
  \item[Exposure] — Heat, attention, or narrative stress.
  \item[Harm] — Injury or trauma.
\end{description}

\textbf{States:}
\begin{itemize}
  \item \textbf{Maintained} — Reliable and ready.
  \item \textbf{Neglected} — Needs downtime or care.
  \item \textbf{Compromised} — Captured, defected, or lost.
\end{itemize}

\subsection*{Upkeep}
\index{Followers!upkeep}

\begin{itemize}
  \item Each Downtime, pay XP equal to Cap or spend a Scene tending the relationship.
  \item Risk: If the GM spends 2+ Complication Points \index{Complication Point (CP)} on an action you take with assistance, they may endanger, injure, or separate the follower instead of you if fictionally appropriate.
  \item Off-Screen Capability: Once per downtime, a Cap 5 follower can solve one significant problem but generates 1 CP for party. The GM must describe how the follower's action creates story consequences for the CP generated.
\end{itemize}

\section{Follower Initiative Actions}
\index{Followers!initiative actions}

Once per scene (across the party), one on-screen follower may take a small independent action:

\begin{itemize}
  \item Scout \& Signal — Change an ally's next action position to Controlled.
  \item Distract \& Draw — Reduce a kinetic rail (Hunt/Escape/Hazard) by –1 tick.
  \item Fetch \& Carry — Move a small object through danger.
\end{itemize}

\textbf{Cost:} Mark Exposure +1 or Harm 1 on that follower.

\section{Loyalty \& Bonds (Optional)}
\index{Followers!loyalty}

\begin{itemize}
  \item Track a simple Loyalty tag per follower: Wary / Steady / Devoted.
  \item Devoted followers can once per arc convert one GM Complication targeting them into a lesser setback.
  \item Wary followers cost +1 XP to Maintain.
\end{itemize}

\section{Summary}

Assets and Followers are powerful tools—but they come with obligations:

\begin{itemize}
  \item Assets solve problems off-screen but require maintenance.
  \item Followers assist in scenes but are vulnerable to Complications.
  \item Both shape the story and reflect your character's influence.
\end{itemize}

Invest wisely—and be ready to pay the price.

\end{chapter}
