% !TEX root = ../fates_edge_players_guide.tex

\chapter{Example Character Builds}
\label{ch:example-builds}

In \textbf{Fate's Edge}, character creation is a blend of mechanical choices and narrative vision. This chapter presents ten fully-realized example characters, each representing a different archetype, culture, and advancement path. Use them as pregens, inspiration, or benchmarks for your own creations.

\section{How to Read These Builds}
\index{Character builds!reading}

Each character includes:
\begin{itemize}
  \item \textbf{Concept} — Narrative identity and role.
  \item \textbf{Build Summary} — Attributes, Skills, Assets, and Followers spent.
  \item \textbf{Dice Pools} — Common rolls and their effectiveness.
  \item \textbf{Signature Move} — Unique approach or ability.
  \item \textbf{Complication Hooks} — Where the GM can apply pressure.
  \item \textbf{Advancement Path} — Suggested next steps for growth.
\end{itemize}

All builds use standard XP costs: Attributes (new × 3), Skills (new × 2), Followers (Cap²), Assets by tier.

\section{1. The Oath-Bound Blade (Paladin)}
\index{Character builds!Oath-Bound Blade}

\textbf{Concept:} A holy knight whose power is tied to an unbreakable vow.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Body 4 (12 XP), Spirit 3 (9 XP)
  \item Melee 4 (2+4+6+8 = 20 XP)
  \item Talent: Oath-Keeper's Word (6 XP)
\end{itemize}
\textbf{Total:} 47 XP (Start at 47 XP tables or adjust)

\textbf{Dice Pools:}
\begin{itemize}
  \item Body + Melee: 8 dice (brutal combat)
  \item Spirit + Melee: 7 dice (focused strikes)
  \item Spirit + Resolve: 6 dice (resist temptation)
\end{itemize}

\textbf{Signature Move:} "By My Sacred Oath" — When protecting an innocent, start all defensive actions Controlled.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Breaking the oath brings spiritual backlash.
  \item Healing transfers harm to the caster.
  \item Binding evil leaves lasting scars.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Presence 2→3 (9 XP) for better leadership
  \item Long term: Spirit-Shield Talent (15 XP) to protect allies
\end{itemize}

\section{2. The Death-Speaker (Necromancer)}
\index{Character builds!Death-Speaker}

\textbf{Concept:} A master who bargains with the dead rather than commands them.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Spirit 4 (12 XP), Wits 3 (9 XP)
  \item Arcana 3 (2+4+6 = 12 XP), Sway 2 (2+4 = 6 XP)
  \item Asset: Minor Shrine to Forgotten Gods (4 XP)
\end{itemize}
\textbf{Total:} 43 XP

\textbf{Dice Pools:}
\begin{itemize}
  \item Wits + Arcana: 6 dice (ritual casting)
  \item Spirit + Sway: 6 dice (negotiating with spirits)
  \item Spirit + Resolve: 7 dice (resisting spiritual corruption)
\end{itemize}

\textbf{Signature Move:} "Whisper to the Departed" — Once per scene, gain insight from a willing spirit.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Summoning requires Boons or accepting Complications.
  \item Undead army is a Major Asset that can revolt if neglected.
  \item Local clergy views you as an abomination.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Spirit 4→5 (15 XP) for greater resilience
  \item Long term: Deathless Servant Asset (12 XP) or Warghost Familiar (9 XP)
\end{itemize}

\section{3. The Border-Warden (Ranger)}
\index{Character builds!Border-Warden}

\textbf{Concept:} A sworn custodian of a liminal territory.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Wits 4 (12 XP), Body 3 (9 XP)
  \item Survival 3 (2+4+6 = 12 XP), Stealth 2 (2+4 = 6 XP)
  \item Asset: Lodge (Minor, 4 XP)
  \item Follower: Scout (Cap 3, 9 XP)
\end{itemize}
\textbf{Total:} 50 XP (Start at 50 XP tables or adjust)

\textbf{Dice Pools:}
\begin{itemize}
  \item Wits + Survival: 7 dice (tracking, navigation)
  \item Body + Stealth: 5 dice (ambush, escape)
  \item Scout Assist: Up to +3 dice on movement/stealth
\end{itemize}

\textbf{Signature Move:} "Territory's Eye" — In your domain, all perception rolls gain +1 effect.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Power is limited geographically.
  \item Every choice risks the balance of warded lands.
  \item Local spirits demand tribute for passage.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Wits 4→5 (15 XP) for greater perception
  \item Long term: Valewood Grove Sanctuary (12 XP) or Beast-Tongue Talent (8 XP)
\end{itemize}

\section{4. The Guild-Approved Shadow (Rogue)}
\index{Character builds!Guild-Approved Shadow}

\textbf{Concept:} A licensed operative of a Guild or Noble's Network.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Wits 4 (12 XP), Presence 3 (9 XP)
  \item Skulduggery 3 (2+4+6 = 12 XP), Subterfuge 2 (2+4 = 6 XP)
  \item Asset: Guild Safehouse (Minor, 4 XP)
  \item Talent: Guild Ties (5 XP)
\end{itemize}
\textbf{Total:} 44 XP

\textbf{Dice Pools:}
\begin{itemize}
  \item Wits + Skulduggery: 7 dice (locks, traps, dirty tricks)
  \item Presence + Subterfuge: 5 dice (deception, social manipulation)
  \item Intricate Actions: Re-roll all 1s for social/engineering rolls
\end{itemize}

\textbf{Signature Move:} "Guild Sanction" — Once per session, invoke guild authority for start Controlled.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Calling on Guild resources costs Boons/XP.
  \item Failure Compromises the Guild Asset.
  \item Rival guilds seek to expose your methods.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Skulduggery 3→4 (8 XP) for master-level skill
  \item Long term: Spy Ring Asset (8 XP) or Shadowbinder Talent (16 XP)
\end{itemize}

\section{5. The Spirit-Touched Outlander (Barbarian)}
\index{Character builds!Spirit-Touched Outlander}

\textbf{Concept:} Rage is trance-possession by an ancestral spirit.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Body 4 (12 XP), Spirit 3 (9 XP)
  \item Melee 3 (2+4+6 = 12 XP), Survival 2 (2+4 = 6 XP)
  \item Talent: Blood Memory (7 XP)
  \item Background: Ykrul Moor Warden (includes Talent)
\end{itemize}
\textbf{Total:} 46 XP

\textbf{Dice Pools:}
\begin{itemize}
  \item Body + Melee: 7 dice (brutal combat)
  \item Spirit + Melee: 6 dice (focused strikes)
  \item Body + Survival: 6 dice (endurance, tracking)
\end{itemize}

\textbf{Signature Move:} "Ancestral Fury" — Once per combat, gain +3 dice but risk Complications.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Gain +3 dice but risk friend/foe confusion.
  \item Compulsive pursuit of worthy opponents.
  \item Rage ends with Fatigue and social isolation.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Body 4→5 (15 XP) for legendary strength
  \item Long term: Warglord Talent (18 XP) or Spirit Shield (15 XP)
\end{itemize}

\section{6. The Scholar of Fractured Truths (Wizard)}
\index{Character builds!Scholar of Fractured Truths}

\textbf{Concept:} Power comes from contradictory fragments of lore.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Wits 4 (12 XP), Spirit 3 (9 XP)
  \item Lore 3 (2+4+6 = 12 XP), Arcana 3 (2+4+6 = 12 XP)
  \item Asset: Grimoire (Minor, 4 XP)
  \item Talent: Ritual Master (12 XP)
\end{itemize}
\textbf{Total:} 53 XP (Start at 53 XP tables or adjust)

\textbf{Dice Pools:}
\begin{itemize}
  \item Wits + Lore: 7 dice (research, ancient knowledge)
  \item Wits + Arcana: 7 dice (spellcasting)
  \item Spirit + Research: 6 dice (deep study)
\end{itemize}

\textbf{Signature Move:} "Fractured Insight" — Once per session, combine two incompatible magical theories for unique effect.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Each "spell" is a volatile truth.
  \item Activation requires Boons/XP.
  \item Grimoire Asset may be Compromised by theft or misuse.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Arcana 3→4 (8 XP) for greater magical power
  \item Long term: Grand Forge-Citadel (20 XP) or Conduit Walker Talent (15 XP)
\end{itemize}

\section{7. The Caretaker of Cycles (Druid)}
\index{Character builds!Caretaker of Cycles}

\textbf{Concept:} A custodian of life-and-death balance.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Spirit 4 (12 XP), Wits 3 (9 XP)
  \item Survival 3 (2+4+6 = 12 XP), Lore 2 (2+4 = 6 XP)
  \item Asset: Grove (Minor, 4 XP)
  \item Talent: Rune-Keeper (8 XP)
\end{itemize}
\textbf{Total:} 49 XP

\textbf{Dice Pools:}
\begin{itemize}
  \item Spirit + Survival: 7 dice (nature connection)
  \item Wits + Lore: 5 dice (natural knowledge)
  \item Spirit + Resolve: 7 dice (resist corruption)
\end{itemize}

\textbf{Signature Move:} "Balance's Price" — Healing demands equal decay; choose the cost.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Using power creates counterweight Complications.
  \item Grove Asset reflects successes or failures.
  \item Seasonal changes affect abilities.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Spirit 4→5 (15 XP) for greater divine connection
  \item Long term: Valewood Grove Sanctuary (12 XP) or Beast-Tongue Talent (8 XP)
\end{itemize}

\section{8. The Chronicler of Consequences (Bard)}
\index{Character builds!Chronicler of Consequences}

\textbf{Concept:} A living archive whose songs define history itself.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Presence 4 (12 XP), Wits 3 (9 XP)
  \item Performance 3 (2+4+6 = 12 XP), Insight 2 (2+4 = 6 XP)
  \item Talent: Echo-Courtier (8 XP)
  \item Asset: Performance Venue (Minor, 6 XP)
\end{itemize}
\textbf{Total:} 51 XP

\textbf{Dice Pools:}
\begin{itemize}
  \item Presence + Performance: 7 dice (inspiration, leadership)
  \item Wits + Insight: 5 dice (reading motives, social nuance)
  \item Presence + Sway: 6 dice (negotiation, persuasion)
\end{itemize}

\textbf{Signature Move:} "Song of Consequence" — Once per session, reframe a social situation through performance.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Inspiration weaves allies into a story.
  \item Failure twists it into tragedy.
  \item Patronage and audiences are fragile Assets.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Performance 3→4 (8 XP) for master-level artistry
  \item Long term: University College (12 XP) or Whisper-Lord Talent (8 XP)
\end{itemize}

\section{9. The Ascetic of the Unbound Body (Monk)}
\index{Character builds!Ascetic of the Unbound Body}

\textbf{Concept:} Detachment unbinds body and spirit.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Body 3 (9 XP), Spirit 4 (12 XP)
  \item Melee 2 (2+4 = 6 XP), Meditation 3 (2+4+6 = 12 XP)
  \item Talent: Battle Chant (8 XP)
  \item Background: Sepulcher Adept (includes Talent)
\end{itemize}
\textbf{Total:} 51 XP

\textbf{Dice Pools:}
\begin{itemize}
  \item Body + Melee: 5 dice (unarmed combat)
  \item Spirit + Meditation: 7 dice (resist harm, focus)
  \item Spirit + Resolve: 7 dice (mental fortitude)
\end{itemize}

\textbf{Signature Move:} "Flowing Stillness" — Ignore first Fatigue/Harm each scene by spending Boon.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Harm returns later as delayed Complications.
  \item Monastic Assets impose obligations and alienation.
  \item Physical detachment attracts unwanted attention.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Spirit 4→5 (15 XP) for legendary resilience
  \item Long term: Resonant Architect Talent (18 XP) or Healing Sanctuary (8 XP)
\end{itemize}

\section{10. The Petitioner of a Silent God (Cleric)}
\index{Character builds!Petitioner of a Silent God}

\textbf{Concept:} Miracles are interpretations of divine silence.

\textbf{Build Summary (30 XP):}
\begin{itemize}
  \item Spirit 4 (12 XP), Presence 3 (9 XP)
  \item Faith 3 (2+4+6 = 12 XP), Sway 2 (2+4 = 6 XP)
  \item Asset: Minor Shrine (4 XP)
  \item Talent: Everflame Keeper (16 XP)
\end{itemize}
\textbf{Total:} 51 XP

\textbf{Dice Pools:}
\begin{itemize}
  \item Spirit + Faith: 7 dice (divine power)
  \item Presence + Sway: 5 dice (negotiation, preaching)
  \item Spirit + Resolve: 7 dice (resist temptation)
\end{itemize}

\textbf{Signature Move:} "Silent Mandate" — Once per session, divine will speaks through apparent coincidence.

\textbf{Complication Hooks:}
\begin{itemize}
  \item Each divine act requires justification.
  \item If logic falters, GM twists outcomes.
  \item Temple Asset demands obedience and testing.
\end{itemize}

\textbf{Advancement Path:}
\begin{itemize}
  \item Short term: Faith 3→4 (8 XP) for greater miracles
  \item Long term: Temple Complex (12 XP) or Warden of Storms Talent (18 XP)
\end{itemize}

\section{Using These Builds}
\index{Character builds!usage}

These characters are designed to be:
\begin{itemize}
  \item \textbf{Immediate Play:} Ready for any Fate's Edge game.
  \item \textbf{Flexible:} Easy to adjust for different XP starting points.
  \item \textbf{Inspiring:} Show different paths through the XP system.
  \item \textbf{Balanced:} Each has strengths, weaknesses, and clear growth paths.
\end{itemize}

Whether you're new to Fate's Edge or looking for fresh inspiration, these builds offer a window into the rich possibilities of character creation in a world of consequences.

\end{chapter}
