% !TEX root = ../fates_edge_players_guide.tex

\chapter{Cultures and Peoples}

The world of \textbf{Fate’s Edge} is home to a rich tapestry of cultures, each with its own traditions, values, and unique \textbf{Affinities}—narrative edges that reflect their place in the world. This chapter introduces the major peoples and cultures you may encounter or belong to.

\section{What Is an Affinity?}

An \textbf{Affinity} is a cultural or racial trait that provides a narrative edge rather than a mechanical bonus. It represents how the world itself leans in your favor—whether through ancient pacts, mystical heritage, or deep understanding of the land.

Affinities often unlock access to specific Talents, Arts, or social reactions.

\section{Major Cultures and Peoples}

\subsection*{Humans}

\textbf{Affinity:} Versatility \\
\textbf{Themes:} Adaptability, ambition, diversity

Humans are found across every province and sea. Adaptable and ambitious, they thrive in any environment and quickly adopt the customs of others. Their strength lies not in any single tradition, but in their ability to blend and innovate.

\textbf{Notable Regions:} Vililan, Akilan, Ostrilan \\
\textbf{Example Talents:} Versatile, Guild Ties, Banner-Bearer

\subsection*{Dwarves (Aeler)}

\textbf{Affinity:} Stone-Sense \\
\textbf{Themes:} Craft, tradition, collective memory

Dwarves are children of stone and forge. They value craftsmanship, kinship, and the deep laws of the earth. Their cities are marvels of engineering, and their word is as unyielding as the mountains they call home.

\textbf{Sub-Types:}
\begin{itemize}
  \item Mountain Dwarves — Deep kingdom dwellers; focused on giants and dragons.
  \item Hill Dwarves — Surface allies; fight Ykrul and trolls.
  \item True Masons — Itinerant stone-clerics who maintain ancient infrastructure.
\end{itemize}

\textbf{Example Talents:} Stone-Sense, Spirit Shield, Forge-Patriarch

\subsection*{Elves (Lethai)}

\textbf{Affinity:} Varies by sub-type \\
\textbf{Themes:} Memory, nature, continuity

Elves see themselves as the first people, divided by philosophy and time. They are long-lived and deeply connected to the natural and mystical world.

\textbf{Sub-Types:}
\begin{itemize}
  \item \textbf{Wood Elves (Lethai-al)} — Tribal and fey; cycle-focused. \\
    Affinity: Backlash Soothing
  \item \textbf{High Elves (Lethai-thora)} — Sequestered exiles of the mind; memory-keepers. \\
    Affinity: Lorekeeper
  \item \textbf{Grass/Dusk/Sundered Elves} — Nomadic high elves estranged from both cousins.
\end{itemize}

\textbf{Example Talents:} Backlash Soothing, Ranger's Step, Echo-Walker

\subsection*{Ykrul}

\textbf{Affinity:} Blood Memory \\
\textbf{Themes:} War, pragmatism, kinship

The Ykrul are a people of horse, boar, and warg. They value strength, loyalty, and the wisdom of ancestors. Their warbands are feared across the steppes, but their hearths are warm and their oaths are binding.

\textbf{Sub-Types:}
\begin{itemize}
  \item Steppe Ykrul — Pastoral nomads of the Violet Steppes.
  \item Mountain Ykrul — Stockier clans of the highlands.
  \item River Ykrul — Wargboat raiders and coastal plunderers.
\end{itemize}

\textbf{Example Talents:} Blood Memory, Blood Frenzy, Warglord

\subsection*{Gnomes (Aelinnel)}

\textbf{Affinity:} Numerical Insight \\
\textbf{Themes:} Mathematics, craft, curiosity

Gnomes are known for their incomprehensible mathematics and delicate craft. They are small in stature but vast in intellect, often found in groves, towers, or burrows deep in the earth.

\textbf{Sub-Types:}
\begin{itemize}
  \item Rock Gnomes — Gemstone engineers of Wrentfell.
  \item Forest Gnomes — Attuned to nature and fey magic.
  \item Faetae — Rare, fragile winged gnomes with insectile wings.
\end{itemize}

\textbf{Example Talents:} Numerical Insight, Grove Cantor, Winged Messenger

\subsection*{Halflings (Aelaerem)}

\textbf{Affinity:} Hearth Sense \\
\textbf{Themes:} Community, resilience, pastoral life

Halflings are a pastoral people who value hearth, home, and harmony with the land. They are resilient, friendly, and often underestimated—until their slings find their mark.

\textbf{Sub-Types:}
\begin{itemize}
  \item Hill Halflings — Farmers and traders of Amedell.
  \item River Halflings — Bargefolk and ferry-keepers.
  \item Wander Halflings — Caravaners and adventurers.
\end{itemize}

\textbf{Example Talents:} Hearth Sense, River Guide, Wanderer's Luck

\subsection*{Tulkani}

\textbf{Affinity:} Shadowbinding \\
\textbf{Themes:} Performance, trade, secrecy

The Tulkani are wanderers, performers, and traders who move like shadows across the northern lands. They are known for their wit, their artistry, and their whispered pacts with powers best left unnamed.

\textbf{Example Talents:} Shadowbinding, Performance Mastery, Guild Ties

\subsection*{Other Peoples}

\begin{description}
  \item[Fharan] — Desert caravans and incense kingdoms. Masters of trade and star-watching.
  \item[Sidhi] — Coastal and river peoples of Akilan. Builders, scribes, and traders.
  \item[Kuvani] — Steppe riders of Dhahara. Honor-bound and fiercely independent.
  \item[Ashaani] — Ancient silt kingdoms. Masters of necropolises and divine law.
\end{description}

\section{Mixed Heritage}

In Fate’s Edge, mixed heritage is common and celebrated. Characters may claim multiple cultural Affinities, but must choose which one to use for Talent prerequisites and cultural benefits.

\section{Summary}

Culture in Fate’s Edge shapes more than appearance—it defines how the world reacts to you:

\begin{itemize}
  \item Affinities provide narrative edges and unlock Talents.
  \item Each culture has its own values, conflicts, and stories.
  \item Heritage can be a source of strength—or a chain to the past.
\end{itemize}

Choose your roots carefully—and let them guide your legend.
