% !TEX root = ../fates_edge_players_guide.tex

\chapter{Attributes and Skills}
\label{ch:attributes-skills}

In \textbf{Fate's Edge}, your character is defined not by a long list of numbers—but by the interplay of four core \textbf{Attributes} and a focused set of \textbf{Skills}. Together, they shape your dice pools, guide your narrative choices, and reflect who you are in the world.

\section{Attributes}
\index{Attribute}

Attributes represent broad aspects of your character's being. Each is rated from 1 to 5 for most mortals. Exceptional beings may exceed this range.

\begin{description}
  \item[Body] \index{Attribute!Body} — Strength, endurance, and physical force. \\
    Used for: Melee, Athletics, Endurance.

  \item[Wits] \index{Attribute!Wits} — Perception, cleverness, and reaction speed. \\
    Used for: Arcana, Stealth, Insight, Tactics.

  \item[Spirit] \index{Attribute!Spirit} — Willpower, intuition, and resilience. \\
    Used for: Meditation, Resolve, Faith, Endurance.

  \item[Presence] \index{Attribute!Presence} — Charm, command, and social force. \\
    Used for: Sway, Command, Performance, Sway.
\end{description}

\section{Skills}
\index{Skill}

Skills are focused expressions of talent. Each is tied to one or more Attributes, but creative players may justify alternative pairings if the narrative fits.

\subsection*{Skill Ratings}

\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Rating} & \textbf{Meaning} \\
\midrule
0 & Untrained — Rely on raw Attribute alone \\
1 & Familiar — Basic competence, a journeyman's touch \\
2 & Skilled — Trained and reliable in most circumstances \\
3 & Expert — Professional mastery or long experience \\
4 & Master — Renowned; your work is widely recognized \\
5 & Legendary — Near-supernatural; few mortals reach this level \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Common Skills}

Below is a list of common skills used in Fate's Edge:

\begin{description}
  \item[Athletics] \index{Skill!Athletics} — Climbing, running, swimming.
  \item[Arcana] \index{Skill!Arcana} — Magical theory, rituals, spellwork.
  \item[Brawl] \index{Skill!Brawl} — Fists, grappling, improvised fighting.
  \item[Insight] \index{Skill!Insight} — Intuition, empathy, lie detection.
  \item[Melee] \index{Skill!Melee} — Blades, axes, polearms.
  \item[Ranged] \index{Skill!Ranged} — Bows, crossbows, thrown arms.
  \item[Sway] \index{Skill!Sway} — Negotiation, mediation, etiquette.
  \item[Stealth] \index{Skill!Stealth} — Hiding, shadowing, evading.
  \item[Deception] \index{Skill!Deception} — Disguise, misdirection, bluffing.
  \item[Survival] \index{Skill!Survival} — Tracking, foraging, navigation.
  \item[Command] \index{Skill!Command} — Leadership, intimidation, rallying.
  \item[Craft] \index{Skill!Craft} — Smithing, alchemy, tinkering.
  \item[Performance] \index{Skill!Performance} — Music, oratory, storytelling.
  \item[Lore] \index{Skill!Lore} — History, cultures, languages.
\end{description}

\section{Building Your Dice Pool}
\index{Dice pool}

When attempting a significant action, combine the relevant Attribute + Skill to determine your dice pool.

\textbf{Examples:}
\begin{itemize}
  \item Body + Melee = raw steel and sinew.
  \item Wits + Melee = anticipation and precision.
  \item Presence + Melee = flourish, feint, and intimidation.
\end{itemize}

The same Skill, seen through different Attributes, tells a different story.

\section{Skill Caps and Growth}
\index{Skill!growth}

Each Skill is capped by its linked Attribute. You cannot raise a Skill above your Attribute rating.

\textbf{Cost to Raise Skills:} New level × 2 XP

\textbf{Example:} Raising Stealth from 2 to 3 costs 3 × 2 = \textbf{6 XP}.

\section{Skill Specializations}
\index{Skill!specialization}

Some skills may be used in specialized contexts. Players may describe niche expertise (e.g., "Desert Survival" or "City Stealth") to gain narrative advantages, subject to GM approval.

\chapter{Additional Skills for Fate's Edge}
\label{ch:additional-skills}

\section{Combat and Tactical Skills}

\begin{skillbox}{Tactics (Wits or Presence)}
The art of battlefield command, strategic planning, and tactical coordination.

\textbf{Applications:} Commanding units, planning ambushes, reading enemy formations, coordinating complex maneuvers.

\textbf{Special Effects:} 
\begin{itemize}
    \item At level 3+: Once per battle, grant all allies +1 die on one coordinated action
    \item At level 5+: Convert one Club complication into a tactical advantage for your forces
\end{itemize}

\textbf{Deck Hooks:} In military regions, convert war-related Clubs into strategic opportunities.
\end{skillbox}

\begin{skillbox}{Weaponcraft (Wits + Body)}
Specialized knowledge of weapons construction, maintenance, and combat techniques.

\textbf{Applications:} Crafting and repairing weapons, identifying weapon qualities, specialized combat techniques.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Identify weapon origin and quality automatically
    \item At level 4+: Modify weapons for specific purposes (+1 effect on one type)
\end{itemize}

\textbf{Deck Hooks:} In crafting regions, reduce Diamond costs for weapon-related leverage.
\end{skillbox}

\begin{skillbox}{Armorcraft (Wits + Body)}
Knowledge of armor construction, maintenance, and defensive techniques.

\textbf{Applications:} Crafting and repairing armor, identifying armor qualities, defensive combat techniques.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Reduce harm from one attack by one level per scene
    \item At level 4+: Create improvised armor from available materials
\end{itemize}

\textbf{Deck Hooks:} Convert physical Clubs (gear damage) into armor modification opportunities.
\end{skillbox}

\section{Social and Cultural Skills}

\begin{skillbox}{Courtly Etiquette (Presence)}
Knowledge of formal social protocols, court behavior, and noble customs.

\textbf{Applications:} Navigating noble courts, formal diplomacy, social events, protocol adherence.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Gain +1 die on all social rolls in formal court settings
    \item At level 5+: Once per session, turn a social faux pas into a charming anecdote
\end{itemize}

\textbf{Deck Hooks:} In noble regions, convert Hearts complications into social advantages.
\end{skillbox}

\begin{skillbox}{Streetwise (Wits)}
Knowledge of urban underworlds, criminal networks, and informal economies.

\textbf{Applications:} Finding black market goods, navigating criminal elements, urban survival.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Know three useful contacts in any major city
    \item At level 4+: Once per session, find exactly what you need on the black market
\end{itemize}

\textbf{Deck Hooks:} Convert Diamond complications into underground network advantages.
\end{skillbox}

\begin{skillbox}{Linguistics (Wits)}
Study of languages, translation, and communication across cultures.

\textbf{Applications:} Learning new languages, translation, code-breaking, linguistic analysis.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Learn basic communication in any language within one day
    \item At level 5+: Create and teach simple pidgin languages for multi-cultural groups
\end{itemize}

\textbf{Deck Hooks:} In multi-cultural regions, downgrade language barrier Clubs to minor delays.
\end{skillbox}

\section{Craft and Trade Skills}

\begin{skillbox}{Engineering (Wits)}
Design and construction of complex structures, machines, and systems.

\textbf{Applications:} Building construction, siege engines, mechanical devices, infrastructure projects.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Design structures that are +1 die more resilient to damage
    \item At level 5+: Create innovative solutions to engineering problems under pressure
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs related to structural failures into engineering challenges.
\end{skillbox}

\begin{skillbox}{Alchemy (Wits + Spirit)}
Creation of potions, elixirs, and chemical substances through mystical means.

\textbf{Applications:} Potion brewing, poison creation, material transmutation, experimental chemistry.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Identify alchemical substances automatically
    \item At level 4+: Create alchemical items with reduced material costs
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (failed experiments) into new recipe discoveries.
\end{skillbox}

\begin{skillbox}{Mercantile (Presence + Wits)}
Expertise in trade, economics, and commercial negotiation.

\textbf{Applications:} Trade negotiation, market analysis, supply chain management, investment.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Gain +1 die on all trading rolls; find better deals automatically
    \item At level 5+: Establish trade relationships that provide ongoing benefits
\end{itemize}

\textbf{Deck Hooks:} Convert economic Clubs into market opportunity discoveries.
\end{skillbox}

\section{Exploration and Survival Skills}

\begin{skillbox}{Navigation (Wits)}
Finding one's way through various terrains and conditions.

\textbf{Applications:} Land, sea, and air navigation; map reading; route planning; wayfinding.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Never get truly lost in familiar terrain
    \item At level 4+: Create accurate maps from memory and observation
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (getting lost) into scenic route discoveries.
\end{skillbox}

\begin{skillbox}{Survival (Wits + Body)}
Living off the land and enduring harsh conditions.

\textbf{Applications:} Finding food and shelter, weather prediction, wilderness first aid, tracking.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Ignore first level of environmental Fatigue in wilderness settings
    \item At level 5+: Thrive in conditions that would challenge others
\end{itemize}

\textbf{Deck Hooks:} Convert environmental Clubs into survival opportunity advantages.
\end{skillbox}

\begin{skillbox}{Tracking (Wits)}
Following trails, reading signs, and understanding movement patterns.

\textbf{Applications:} Pursuing targets, wildlife tracking, forensic investigation, path reading.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Follow any trail regardless of age or conditions
    \item At level 4+: Predict target behavior based on tracking evidence
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (losing the trail) into environmental clue discoveries.
\end{skillbox}

\section{Lore and Knowledge Skills}

\begin{skillbox}{Arcane Lore (Wits + Spirit)}
Study of magical theory, spell construction, and mystical phenomena.

\textbf{Applications:} Spell research, magical item identification, theoretical magic, arcane history.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Reduce magical Backlash by 1 point when researching spells
    \item At level 5+: Identify and counteract magical effects without rolling
\end{itemize}

\textbf{Deck Hooks:} Convert magical Clubs into research breakthrough opportunities.
\end{skillbox}

\begin{skillbox}{Historical Knowledge (Wits)}
Understanding of past events, cultures, and their impacts on the present.

\textbf{Applications:} Historical research, cultural analysis, precedent identification, ancient language.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Recognize historical patterns in current events
    \item At level 4+: Recall specific historical details relevant to current situations
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (historical confusion) into precedent-based solutions.
\end{skillbox}

\begin{skillbox}{Natural Philosophy (Wits)}
Study of the natural world, scientific principles, and empirical observation.

\textbf{Applications:} Scientific observation, hypothesis testing, natural phenomenon explanation.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Make accurate predictions about natural phenomena
    \item At level 5+: Develop innovative solutions to natural world challenges
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (natural disasters) into scientific understanding opportunities.
\end{skillbox}

\section{Stealth and Infiltration Skills}

\begin{skillbox}{Stealth (Wits + Body)}
Moving unseen and unheard through various environments.

\textbf{Applications:} Sneaking, hiding, silent movement, misdirection.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Move silently in most environments without rolling
    \item At level 4+: Create distractions that misdirect attention automatically
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (being detected) into misdirection opportunities.
\end{skillbox}

\begin{skillbox}{Lockpicking (Wits + Body)}
Opening locks and secured mechanisms without keys.

\textbf{Applications:} Lock picking, trap disarming, security system bypass, mechanical puzzle solving.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Identify lock complexity and security level automatically
    \item At level 4+: Create specialized tools from available materials
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (lock failure) into security system understanding.
\end{skillbox}

\begin{skillbox}{Disguise (Presence + Wits)}
Creating and maintaining false identities and appearances.

\textbf{Applications:} Identity creation, costume design, behavioral mimicry, social infiltration.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Maintain simple disguises without rolling
    \item At level 5+: Create complex identities that withstand scrutiny
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (disguise failure) into character development opportunities.
\end{skillbox}

\section{Healing and Medicine Skills}

\begin{skillbox}{Medicine (Wits + Spirit)}
Healing injuries, treating diseases, and understanding anatomy.

\textbf{Applications:} Wound treatment, disease diagnosis, surgical procedures, pharmaceutical preparation.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Stabilize dying patients automatically
    \item At level 4+: Reduce recovery time for treated injuries by 50\%
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (medical complications) into treatment innovation opportunities.
\end{skillbox}

\begin{skillbox}{Herbalism (Wits + Spirit)}
Knowledge of medicinal and poisonous plants and their applications.

\textbf{Applications:} Plant identification, remedy preparation, poison creation, natural healing.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 2+: Identify medicinal properties of plants automatically
    \item At level 4+: Create enhanced remedies with additional beneficial effects
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (plant misidentification) into new botanical discoveries.
\end{skillbox}

\begin{skillbox}{Psychology (Wits + Spirit)}
Understanding of mental states, behavioral patterns, and psychological healing.

\textbf{Applications:} Mental health treatment, behavioral analysis, trauma therapy, mind reading.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Calm panicked individuals automatically in most situations
    \item At level 5+: Help others overcome psychological traumas and fears
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (mental breakdowns) into therapeutic breakthrough opportunities.
\end{skillbox}

\section{Regional and Cultural Skills}

\begin{skillbox}{Utaran Law \& Precedent (Wits or Presence)}
Imperial codes, edicts, and forum custom from Ecktoria to the marches.

\textbf{Applications:} Legal proceedings, bureaucratic navigation, precedent research, contract drafting.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Downgrade one legal issue in this leg to Bureaucratic Delay
    \item At level 5+: Mint a temporary Stay of Writ (Debt Flip: a clerk demands a future favor)
\end{itemize}

\textbf{Deck Hooks:} In Vhasia/Viterra ports, convert quarantine/levy Clubs to inspection (softer bite).
\end{skillbox}

\begin{skillbox}{Stone-Tongue (Spirit)}
Read the memory of stone, unworked rock to vault masonry.

\textbf{Applications:} Archaeological investigation, structural analysis, historical stone reading.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Ask the GM 1-3 questions about who carved/used/strained this stone
    \item At level 5+: Once per Aeler leg, peek at the next leg's travel complications
\end{itemize}

\textbf{Deck Hooks:} Convert stone-related Clubs into historical revelation opportunities.
\end{skillbox}

\begin{skillbox}{Monsoon-Reckoning (Wits)}
Seasonal winds and their social/economic knock-on effects.

\textbf{Applications:} Weather prediction, trade route planning, seasonal activity coordination.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Shift a weather complication down one rank or convert to Forecast boon
    \item At level 5+: Predict major weather events up to one week in advance
\end{itemize}

\textbf{Deck Hooks:} Convert weather Clubs into strategic advantage opportunities.
\end{skillbox}

\begin{skillbox}{Heraldry \& Lineage (Presence)}
Houses, devices, marriages, feuds.

\textbf{Applications:} Noble identification, family history research, heraldic symbol interpretation.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Gain Controlled position vs titled NPCs; ask 1 specific House-Truth on success
    \item At level 5+: Create Letters of Lineage pass (acts as a local Warrant)
\end{itemize}

\textbf{Deck Hooks:} Convert noble-related Clubs into alliance opportunity discoveries.
\end{skillbox}

\begin{skillbox}{Geasa \& Oath-Weaving (Spirit)}
Binding promises, taboos, and loopholes.

\textbf{Applications:} Oath creation, taboo navigation, promise binding, loophole exploitation.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Attach a 4-segment geas clock to a willing target; define trigger \& consequence
    \item At level 5+: Create complex oath networks that bind multiple parties
\end{itemize}

\textbf{Deck Hooks:} Theona/Valewood 9s can satisfy or complicate a geas condition.
\end{skillbox}

\section{Specialized Professional Skills}

\begin{skillbox}{Espionage (Wits + Presence)}
Intelligence gathering, covert operations, and spy craft.

\textbf{Applications:} Information gathering, surveillance, counter-intelligence, covert communication.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Establish secure communication channels automatically
    \item At level 5+: Create and maintain deep cover identities
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (exposure) into intelligence windfall opportunities.
\end{skillbox}

\begin{skillbox}{Sway (Presence + Spirit)}
Negotiation, treaty crafting, and international relations.

\textbf{Applications:} Peace negotiations, treaty drafting, conflict resolution, alliance building.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Reduce tension in hostile negotiations automatically
    \item At level 5+: Forge lasting peace agreements between former enemies
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (diplomatic failures) into creative solution opportunities.
\end{skillbox}

\begin{skillbox}{Investigation (Wits)}
Crime solving, evidence analysis, and logical deduction.

\textbf{Applications:} Crime scene analysis, suspect interrogation, evidence collection, mystery solving.

\textbf{Special Effects:}
\begin{itemize}
    \item At level 3+: Notice crucial evidence others miss automatically
    \item At level 5+: Solve complex mysteries through logical deduction
\end{itemize}

\textbf{Deck Hooks:} Convert Clubs (misleading evidence) into breakthrough opportunities.
\end{skillbox}

\section*{Skill Progression Guidelines}

\subsection*{Rating 0 - Untrained}
Rely on raw Attribute dice only. No special effects.

\subsection*{Rating 1 - Familiar}
Basic competence in the skill. May attempt related actions with standard difficulty.

\subsection*{Rating 2 - Skilled}
Reliable training. Special Effect 1 unlocked for this skill.

\subsection*{Rating 3 - Expert}
Professional mastery. May train others in the skill. Special Effect 2 unlocked.

\subsection*{Rating 4 - Master}
Renowned in your field. Others seek your expertise. Special Effect 3 unlocked.

\subsection*{Rating 5 - Legendary}
Near-mythic talent. Your name is known in connection with this skill. Special Effect 4 unlocked.

\section*{Skill Synergies}

Many skills work well together:
\begin{itemize}
    \item \textbf{Tactics + Command:} Battlefield leadership
    \item \textbf{Alchemy + Medicine:} Advanced healing remedies
    \item \textbf{Engineering + Survival:} Wilderness shelter construction
    \item \textbf{Linguistics + Heraldry:} Cultural translation and protocol
    \item \textbf{Investigation + Psychology:} Behavioral profiling
    \item \textbf{Navigation + Tracking:} Comprehensive pursuit capabilities
\end{itemize}

These synergies can be reflected in group rolls or combined actions where appropriate.

\section{Summary}

Attributes and Skills are the core of your character's identity:

\begin{itemize}
  \item Attributes define your potential.
  \item Skills reflect your training and focus.
  \item Together, they determine how you act—and how the world responds.
\end{itemize}

Choose wisely. Your capabilities are not just tools—they are expressions of who you are.

\end{chapter}
