% !TEX root = ../fates_edge_players_guide.tex

\chapter{Compendium of Assets}
\label{ch:compendium-assets}

In \textbf{Fate's Edge}, \textbf{Assets} \index{Assets} are off-screen resources that extend your influence beyond personal capability. This chapter provides a comprehensive list of Assets organized by tier, from minor holdings to empire-shaping institutions.

\section{Design Principles}
\index{Assets!design principles}

Assets in Fate's Edge are designed to:

\begin{itemize}
  \item Shape the world between sessions.
  \item Provide narrative leverage rather than direct mechanical bonuses.
  \item Create story hooks and complications through upkeep and neglect.
\end{itemize}

Assets do \textbf{not}:
\begin{itemize}
  \item Act in scenes directly (that's what Followers are for).
  \item Create Diamonds automatically.
  \item Stack identical tags—take the best one.
\end{itemize}

\section{Minor Assets (4 XP)}
\index{Assets!Minor}

Small holdings or tools that provide basic leverage.

\begin{description}
  \item[Signature Weapon] — Named heirloom or masterpiece. Off-Screen: Mitigate gear-break CP 1; recovery scene clears Compromised.
  \item[Scholar's Cell] — Quiet room with texts and desk. Off-Screen: One research opener starts Controlled or +1 effect.
  \item[Safehouse Network] — Hidden bolt-holes in one city. Off-Screen: Clear Hunt −1 between scenes or start Controlled for entry.
  \item[Artisan's Workshop] — Tools + bench for field repairs. Off-Screen: Clear Compromised from one item.
  \item[Courier Network] — Trusty runners and dead-drops. Off-Screen: Deliver message; next parley starts Controlled or Mitigate CP 1.
  \item[Small Farm/Orchard] — Larder and surplus. Off-Screen: Waive minor Supply complication or refresh "consumable" gear tag.
  \item[Local Tavern/Inn] — Neighborhood hub. Off-Screen: First social opener starts Controlled.
  \item[Minor Shrine/Temple] — Hearth for the spirit. Off-Screen: Clear Fatigue 1 or Mitigate first morale CP 1.
  \item[Fishing Boat/Barge] — River/harbor skiff. Off-Screen: Water crossing starts Controlled or ignore first travel Supply loss.
  \item[Herbal Garden] — Beds of restorative greens. Off-Screen: Remove Fatigue 1 from one character.
\end{description}

\section{Standard Assets (8 XP)}
\index{Assets!Standard}

Established holdings with regional influence.

\begin{description}
  \item[Noble Title/Charter] — Recognized rank and privileges. Off-Screen: Secure audience; first petition starts Controlled or Protected Turn.
  \item[Guild Membership] — Standing in a working order. Off-Screen: Procurement/commission starts Controlled or Mitigate CP 1.
  \item[Merchant Stall/Front] — Fixed face in the market. Off-Screen: Free upkeep on one Minor Asset or +1 effect on barter.
  \item[Spy Ring/Informants] — Eyes and ears on retainer. Off-Screen: Targeted inquiry starts Controlled or +1 effect; GM adds social Hook.
  \item[Mercenary Contract] — Reach for hired steel. Off-Screen: Introduce temporary Cap 2 unit for one beat (consumes Initiative).
  \item[Trading Charter] — Right of trade along routes. Off-Screen: At customs/gates, Mitigate first Diamonds/Clubs CP 1 or start Controlled.
  \item[Library Archive] — Stacks and custodian. Off-Screen: Deep topic research starts Controlled and +1 effect.
  \item[Craftsman's Quarter] — Shop + apprentices. Off-Screen: Produce quality piece: clear Compromised or grant +1 effect once.
  \item[Temple Complex] — Choirs, courts, care. Off-Screen: Among faithful, gain Protected Turn or Mitigate first Sanctity/Crowd CP 1.
  \item[Port Authority] — Berths, booms, bells. Off-Screen: Priority handling: Mitigate first Curfew/Crowd CP 1 or loading starts Controlled.
\end{description}

\section{Major Assets (12 XP)}
\index{Assets!Major}

Significant holdings with broad influence.

\begin{description}
  \item[City District License] — You run a ward's levers. Off-Screen: In-district, civic opener starts Controlled or set local rail −1.
  \item[Regional Monopoly] — Exclusive right in a staple. Off-Screen: In market scenes, cancel first Diamonds CP.
  \item[Fortress/Stronghold] — A hard place held by your banner. Off-Screen: In siege/raid, ignore first Spades complication.
  \item[Large Estate/Demesne] — Fields, mills, partys. Off-Screen: No minor Supply complications or free upkeep on one additional Asset.
  \item[Naval Fleet Share] — Oars, sails, signal flags. Off-Screen: Ensure safe passage: maritime travel starts Controlled and Mitigate CP 1.
  \item[University College] — Faculty, quad, seal. Off-Screen: Downtime reduces one Skill advance cost by 1 (min 1).
  \item[Banking Charter] — Letters, ledgers, liens. Off-Screen: In finance scenes, Mitigate first Diamonds CP 1; if clean, +1 effect.
  \item[Diplomatic Enclave] — Flag, immunity, ink. Off-Screen: In foreign courts, ignore first Clubs complication or parley starts Controlled.
  \item[Mine/Quarry Operation] — Raw stone, hard people. Off-Screen: Provide materials: clear Compromised on structure/gear or +1 effect.
  \item[Market Control] — You set the bell. Off-Screen: Declare Surplus/Scarcity for staple; negotiation about it starts Controlled.
\end{description}

\section{Tier III Assets (16–20 XP)}
\index{Assets!Tier III}

Empire-shaping institutions with vast reach.

\begin{description}
  \item[Grand Forge–Citadel] — City-sized heat and hammer. Off-Screen: Produce masterworks; ignore first gear complication or clear batch Compromised.
  \item[Harbor Light Consortium] — Beacons from cape to cape. Off-Screen: Dictate right-of-way: ally's maneuver starts Controlled or 10-ft zone Hazard −1.
  \item[Grand Exchange Charter] — Markets move when you breathe. Off-Screen: In halls, cancel one Diamonds CP; if clean, +1 effect.
  \item[Thepyrgosi Great Library] — The north's memory palace. Off-Screen: Major inquiry starts Controlled and +1 effect; Mitigate first scholarly CP 1.
  \item[Kahfagian Fleet Command] — Signals across sea lanes. Off-Screen: Declare blockade/corridor: hostile exits Escape +1 or friendly passage Controlled.
  \item[Oshiiran Hydraulic Empire] — Water is wealth. Off-Screen: Set Irrigated tag for route; cancel first Supply/shortage Diamonds CP.
  \item[Ashaani Palace Complex] — Keys to office and omen. Off-Screen: Political opener starts Controlled and gains Protected Turn.
  \item[Valewood Grove Sanctuary] — Old roots, older vows. Off-Screen: Wilderness scenes, cancel natural Hazard tick or Mitigate CP 1.
  \item[Dwarven Mountain Hall] — Deep law, deeper craft. Off-Screen: Engineering/crafting opener starts Controlled and +1 effect.
  \item[Ykrul Steppe Confederation] — Ten thousand hooves. Off-Screen: On steppe, group movement starts Controlled and Escape −1.
\end{description}

\section{Specialized Assets}
\index{Assets!specialized}

Assets with unique thematic purposes.

\begin{description}
  \item[Magical Laboratory (10 XP)] — Arcane benches, wards. Off-Screen: Ritual Backlash −1 CP once; experiment starts Controlled.
  \item[Healing Sanctuary (8 XP)] — Beds, herbs, bells. Off-Screen: Treat patient: clear Harm 1 or Fatigue 2 (downtime only).
  \item[Performance Venue (6 XP)] — Stage + crowd. Off-Screen: Build reputation; first Performance opener starts Controlled or +1 effect.
  \item[Observatory/Star Tower (8 XP)] — Ephemerides and lenses. Off-Screen: Predict timing; navigation/timing opener starts Controlled.
  \item[Alchemical Garden (10 XP)] — Beds of rare reagents. Off-Screen: Produce draught; once/session Mitigate first toxin CP 1 or +1 effect salve.
  \item[Training Academy (12 XP)] — Masters, drills, mats. Off-Screen: Downtime reduces one Skill advance cost by 1 (min 1) for one trainee.
  \item[Intelligence Bureau (10 XP)] — Analysts and cut-outs. Off-Screen: Avoid surprise; first recon opener starts Controlled.
  \item[Hospitality House (6 XP)] — Beds, bread, rumor. Off-Screen: Gain Protected Turn in hosted parley or Mitigate first Crowd CP 1.
  \item[Storage Vault (6 XP)] — Locks, ledgers, layers. Off-Screen: Ignore first theft/sabotage complication targeting stored goods.
  \item[Communication Relay (8 XP)] — Fires, flags, drums. Off-Screen: Coordinate afar: timed coordination starts Controlled or Mitigate CP 1.
\end{description}

\section{Asset Upgrades and Variants}
\index{Assets!upgrades}

Enhance your Assets with additional investments:

\begin{description}
  \item[Enhanced Security (+2 XP)] — Hardened doors, better watchers. Effect: Mitigate first theft/sabotage CP 1 each session.
  \item[Expanded Scope (+3 XP)] — Wider footprint. Effect: Use Off-Screen effect twice/session or across two adjacent venues.
  \item[Magical Enhancement (+4 XP)] — Warded, sanctified. Effect: Add Mitigate Backlash 1 (once/session) for rituals OR Hazard −1 aura (10-ft, 1 beat).
  \item[Political Connections (+3 XP)] — Papers, patrons. Effect: In civic/authority scenes, Mitigate first Clubs CP 1.
\end{description}

\section{Summary}

Assets are the tools of influence and legacy:

\begin{itemize}
  \item Minor Assets provide personal leverage.
  \item Standard Assets offer regional reach.
  \item Major Assets shape politics and trade.
  \item Tier III Assets reshape the setting itself.
\end{itemize}

Choose your holdings wisely—and be ready to pay their price.

\end{chapter}
