% !TEX root = ../fates_edge_players_guide.tex

\chapter{The Three XP Paths}
\label{ch:xp-paths}

In \textbf{Fate's Edge}, how you spend your Experience Points (XP) \index{Experience Points (XP)} defines not only your character's capabilities—but also their role in the world. There are three broad paths of advancement, each representing a different philosophy of growth:

\begin{description}
  \item[Enhance Self] \index{XP path!Enhance Self} — Invest in personal mastery.
  \item[Acquire Assets] \index{XP path!Acquire Assets} — Build influence and networks.
  \item[Learn Talents] \index{XP path!Learn Talents} — Unlock unique abilities and cultural arts.
\end{description}

This chapter explores each path, helping you choose the one that fits your vision—and warning you of the risks that come with each.

\section{Path 1: Enhance Self}
\index{XP path!Enhance Self}

The \textbf{Solo Path} focuses on personal mastery. Players who choose this invest the majority of their XP in Attributes and Skills.

\subsection*{Typical XP Spread}
\begin{itemize}
  \item 70–90\% Enhance Self
  \item 0–10\% Assets
  \item 0–20\% Talents
\end{itemize}

\subsection*{Strengths}
\begin{itemize}
  \item Reliable in spotlight scenes.
  \item Few moving parts—less upkeep.
  \item Resilient to follower loss or asset neglect.
\end{itemize}

\subsection*{Risks}
\begin{itemize}
  \item Vulnerable in logistics or social scenes.
  \item May stall when problems require networks or off-screen leverage.
\end{itemize}

\subsection*{Example Build: The Blade-Seeker}
\begin{itemize}
  \item Body 4, Wits 3
  \item Melee 4, Athletics 2
  \item Talents: Battle Instincts, Silver Tongue
\end{itemize}

\section{Path 2: Acquire Assets}
\index{XP path!Acquire Assets}

The \textbf{Mixed Path} balances personal growth with influence. Players invest in both self and assets—followers, holdings, or networks.

\subsection*{Typical XP Spread}
\begin{itemize}
  \item 50–65\% Enhance Self
  \item 15–25\% Assets
  \item 15–25\% Talents
\end{itemize}

\subsection*{Strengths}
\begin{itemize}
  \item Adaptable—strong in many scenes.
  \item Bridges party gaps.
  \item Can solve problems off-screen.
\end{itemize}

\subsection*{Risks}
\begin{itemize}
  \item Upkeep pressure—assets and followers require attention.
  \item Followers can be targeted by GM Complication Points.
\end{itemize}

\subsection*{Example Build: The Bond-Keeper}
\begin{itemize}
  \item Presence 3, Wits 3
  \item Sway 3, Lore 1
  \item Follower: Cap 3 Archivist
  \item Asset: Minor Safehouse
\end{itemize}

\section{Path 3: Learn Talents}
\index{XP path!Learn Talents}

The \textbf{Mastermind Path} prioritizes influence and narrative power. Players invest heavily in followers, assets, and Talents.

\subsection*{Typical XP Spread}
\begin{itemize}
  \item 25–40\% Enhance Self
  \item 35–55\% Assets
  \item 20–40\% Talents
\end{itemize}

\subsection*{Strengths}
\begin{itemize}
  \item Strong strategic reach.
  \item Scene control via assistance and assets.
  \item Excellent at heists, social manipulation, and long-term planning.
\end{itemize}

\subsection*{Risks}
\begin{itemize}
  \item Dependency on lanes and helpers.
  \item Complication fallout can cascade.
  \item Vulnerable if assets are compromised or followers lost.
\end{itemize}

\subsection*{Example Build: The Spider}
\begin{itemize}
  \item Wits 3, Presence 3
  \item Tactics 2, Subterfuge 2
  \item Followers: Cap 4 Scout, Cap 3 Informant
  \item Assets: Standard Spy Network, Minor Workshop
\end{itemize}

\section{Choosing Your Path}

Your XP path is not fixed. You may shift over time, but each path comes with trade-offs:

\begin{center}
\begin{tabular}{lccc}
\toprule
\textbf{Path} & \textbf{Focus} & \textbf{Strength} & \textbf{Fragility} \\
\midrule
Solo & Attributes/Skills & High personal pools & Low (self-reliant) \\
Mixed & Balanced & Moderate pools + situational help & Medium (helper risk) \\
Mastermind & Assets + Followers & Strong leverage & High (network risk) \\
\bottomrule
\end{tabular}
\end{center}

\section{Summary}

Your XP path defines your role at the table:

\begin{itemize}
  \item Will you be the hero who stands alone?
  \item Will you balance strength with support?
  \item Will you command webs of influence and intrigue?
\end{itemize}

Each path offers unique rewards—and consequences. Choose wisely, and let your choices echo through the world.

\end{chapter}
