% !TEX root = ../fates_edge_players_guide.tex

\chapter{Talents}
\label{ch:talents}

In \textbf{Fate's Edge}, \textbf{Talents} \index{Talents} are unique abilities that expand what your character can do. They are purchased with XP and often serve as stepping stones toward Prestige Abilities. Talents can be general, cultural, or narrative capstones that reshape how you engage with the world. Some Talents and Prestige Abilities incorporate \textbf{Tags} (see Chapter \ref{ch:tags}), which are standardized mechanical shorthand for common effects.

\section{Types of Talents}
\index{Talents!types}

Talents are organized into three categories based on cost and impact:

\subsection*{General Talents}
\index{Talents!general}

\begin{itemize}
    \item \textbf{Battle Instincts} (Cost: 6 XP): Once per scene, re-roll a failed defense roll.
    \item \textbf{Silver Tongue} (Cost: 4 XP): Gain +1 die when persuading or deceiving through speech.
    \item \textbf{Iron Stomach} (Cost: 3 XP): Immune to mundane poisons and spoiled food; halve Complications from toxic sources.
    \item \textbf{Exceptional Coordination} (Cost: 8 XP): One follower can provide +4 assist dice (instead of the normal maximum of +3).
    \item \textbf{Danger Sense} (Cost: 5 XP): Once per scene, treat the first Story Beat generated against you as if it were a 1 lower (minimum 1). This can turn a 4 SB spend into a 3 SB spend, for example.
    \item \textbf{Steady Hands} (Cost: 4 XP): Ignore the first penalty die from Harm or Fatigue on any action involving fine motor control (e.g., picking locks, crafting, using ranged weapons).
\end{itemize}

\subsection*{Racial or Cultural Talents}
\index{Talents!racial/cultural}

\begin{itemize}
    \item \textbf{Stone-Sense} (Dwarves, Cost: 5 XP): Detect flaws in stone or earth; gain +1 die on Engineering or Craft rolls underground.
    \item \textbf{Backlash Soothing} (Wood Elves, Cost: 6 XP): Once per session, reduce a magical Backlash Complication by 2 points when in natural terrain.
    \item \textbf{Blood Memory} (Ykrul, Cost: 5 XP): After a battle, meditate to gain one temporary Skill die reflecting a foe's tactics for the next scene.
    \item \textbf{Fae-Touched} (Fae Courts, Cost: 7 XP): Gain +1 die on rolls to resist or interact with Fae magic. Once per session, you may treat a roll of 1-2 on a d10 as a 6 for the purposes of generating a Boon from a failure, but you must accept a minor, immediate, Fae-themed complication (GM's discretion).
    \item \textbf{Deep Law Keeper} (Aeler, Cost: 6 XP): Gain +1 die on rolls related to dwarven law, tradition, or resisting enchantments while underground.
\end{itemize}

\subsection*{Prestige Abilities}
\index{Prestige Abilities}

Prestige Abilities are high-impact Talents, often with significant narrative weight and sometimes incorporating Tags. They frequently have multiple prerequisites.

\begin{itemize}
    \item \textbf{Echo-Walker's Step} (High Elf, Cost: 20 XP; Req: Wits 5, Arcana 4): 
1/arc, \emph{observe} a perfect echo of a past event at your location (no retconning). 
GM immediately banks +2 SB; scenes touching that memory carry an omen. Grants DV −1 on one action that uses the revealed truth.
    \item \textbf{Warglord} (Ykrul, Cost: 18 XP; Req: Body 5, Command 3): 
Once per campaign, unify scattered warbands into a single host for a season. Start a \emph{Logistics} clock and a \emph{Grudge} clock; either one filling fractures the host.
    \item \textbf{Spirit-Shield} (Aeler, Cost: 15 XP; Req: Spirit 4, Insight 3): 
1/session, erase up to 3 SB from an ally's \emph{current} roll; you immediately mark Fatigue +1 and the GM banks +1 SB as backlash.
    \item \textbf{Warder's Seal} (Cost: 15 XP; Req: Wits 4, Arcana 3): You may create a permanent magical \TagName{WARD} (see Chapter \ref{ch:tags}) as a Downtime activity. The DV to cross is equal to your Arcana rating. Crafting the \TagName{WARD} costs 2 XP and requires specific materials. Maintaining multiple \TagName{WARD}s requires XP upkeep (1 XP per \TagName{WARD} per month, paid during Downtime).
    \item \textbf{Theurgy} (Cost: 25 XP; Req: Wits 5, Faith 4, relevant Cleric Talent): You may perform minor miracles aligned with your faith. Once per session, you can invoke an effect equivalent to a \TagName{BARRIER}, \TagName{CLEANSE}, or \TagName{FORTIFY} Tag (see Chapter \ref{ch:tags}) without needing to cast a spell or spend Boons. Define the effect and cost with the GM (typically DV 2-3, scene duration). Performing a miracle generates 1 SB for the GM.
    \item \textbf{Runekeeper's Bond} (Cost: 12 XP; Req: Wits 4, Arcana 3): You may bind a Thiasos (a supernatural familiar) to a specific Patron (see Chapter \ref{ch:gods-powers-patrons}). This is a prerequisite for learning and using Rites associated with that Patron. The Thiasos follows standard Familiar rules for Exposure and Harm.
\end{itemize}

\section{Purchasing Talents}
\index{Talents!purchasing}

Talents are purchased with XP and may have prerequisites:

\begin{itemize}
  \item \textbf{Attribute/Skill Requirements} — Must meet minimum ratings (permanent, not temporary).
  \item \textbf{Asset/Tag/Talent Requirements} — Some Talents require ownership of specific assets, the ability to use certain Tags, or possession of other Talents.
  \item \textbf{Cultural Requirements} — Certain Talents are restricted to specific cultures or Affinities.
  \item \textbf{Downtime} — Spending XP on a Talent usually requires an equivalent amount of Downtime to represent training or attunement.
\end{itemize}

\textbf{Example:} \textit{Warder's Seal} requires Wits 4, Arcana 3, and represents a significant investment of time and resources.

\section{Prestige Prerequisites}
\index{Prestige Abilities!prerequisites}

\begin{itemize}
    \item Qualifying: Attribute/Skill/Talent prerequisites must be met with permanent ratings.
    \item After purchase: If you later lose a prerequisite (e.g., an Attribute is reduced), you \textbf{keep the Talent} but may not be able to use features that require that prerequisite until it is restored.
    \item Scaling: Many Prestige Abilities can be enhanced with additional XP investment, often to increase uses or potency.
\end{itemize}

\section{Using Talents}
\index{Talents!usage}

Talents are typically used once per scene, session, or arc, as noted in their description. Overuse is restricted to prevent imbalance.

\begin{itemize}
  \item \textbf{Once per Scene} — Can be used multiple times per session, but not on the same action.
  \item \textbf{Once per Session} — Limited to one use between downtimes.
  \item \textbf{Once per Arc/Campaign} — Reserved for narrative milestones.
  \item \textbf{Passive/Constant} — Some Talents provide a continuous benefit (e.g., Iron Stomach).
\end{itemize}

\section{Losing Talent Access}
\index{Talents!losing access}

If you later lose a prerequisite (e.g., an Attribute is reduced or an Asset is lost), you \textbf{keep the Talent} but may not be able to use features that require that prerequisite until it is restored.

\section{Summary}

Talents are the unique expressions of your character's growth:

\begin{itemize}
  \item General Talents offer reliable mechanical benefits for common situations.
  \item Cultural Talents reflect your heritage, specialized training, or unique background.
  \item Prestige Abilities reshape the story at key moments, often incorporating powerful effects or Tags.
\end{itemize}

Choose Talents that reflect your character's journey—and let them echo through the world.

\end{chapter}
