% !TEX program = pdflatex
\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{multicol}
\usepackage{titlesec}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=black,citecolor=black}
\setlist{itemsep=2pt,topsep=4pt}
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

\newcommand{\SuitSpade}{\textbf{Spade}} % place
\newcommand{\SuitHeart}{\textbf{Heart}} % actor
\newcommand{\SuitClub}{\textbf{Club}}   % pressure
\newcommand{\SuitDiamond}{\textbf{Diamond}} % leverage
\newcommand{\ClockSizes}{2--5$\Rightarrow$4 • 6--10$\Rightarrow$6 • J/Q/K$\Rightarrow$8 • A$\Rightarrow$10}

\newenvironment{Leg}[1]{\begin{tcolorbox}[enhanced,breakable,sharp corners,boxrule=.6pt, colback=white,colframe=black,title={#1}]}{\end{tcolorbox}}

\begin{document}
\begin{center}
{\LARGE \textbf{Roads Between Realms}}\\[4pt]
{\large A Standalone Travel & Route Reference for Fate's Edge}\par
\vspace{4pt}\small Version 1.0
\end{center}

\tableofcontents
\vspace{6pt}

\section{Scope & Assumptions}
This document connects the regional 52-card generators into a coherent overland, river, sea, pass, and shadow-corridor travel system. It standardizes names, clarifies adjacency, and provides a consistent table procedure for journeys of any length.

\paragraph{Card mapping (standardized).} \SuitSpade{}=\emph{Place}, \SuitHeart{}=\emph{Actor/Faction}, \SuitClub{}=\emph{Complication/Pressure}, \SuitDiamond{}=\emph{Reward/Leverage}. Draw until all four suits appear for a leg; highest rank sets the travel clock (\ClockSizes). \SuitDiamond{} results are \emph{codified outcomes} (permits, charters, escorts, exceptions) that change position rather than call for a roll.

\section{Regional Topology (West \textrightarrow{} East)}
\subsection*{Kahfagia and the Western Reach}
\textbf{Kahfagia} anchors the far west on the Amaranthine. Its convoy law and lantern beacons control the approaches. To its east lies \textbf{Ecktoria}; to the west, open Great Sea routes.

\subsection*{Ecktoria, Acasia, and the Central Peninsula}
\textbf{Ecktoria} occupies a compact peninsula with its self-named capital on the southern/eastern Amaranthine coast. \textbf{Acasia} lies directly to the east. Its sole metropolis is \textbf{Silkstrand}, a deep-water port and regional entrep\^ot on the Amaranthine; even overland caravans risk the broken marches to reach its exchange floors. The Aelerian Mountains touch the northwestern corner of the Ecktoria--Acasia border.

\subsection*{Vhasia and the Inner March}
\textbf{Vhasia} sits east of Acasia. The river \textbf{Belworth} forms the frontier with \textbf{Viterra}. At its mouth stand twin ports: \textbf{Marcott} (Vhasia) and \textbf{Fairport} (Viterra), both opening onto the Amaranthine. North rises the Aelerian range; Aeler-controlled passes lead toward the Mistlands. The \textbf{Lence} flows west/northwest through Vhasian country to meet the Belworth along its mid-reaches.

\subsection*{Viterra and the Twin Mouths}
\textbf{Viterra} stands east of Vhasia. The \textbf{Belworth} marks the shared boundary; at its mouth lie \textbf{Marcott} (Vhasia) and \textbf{Fairport} (Viterra) with access to the Amaranthine. Midway upriver, \textbf{Tarlington} crowns a dramatic plateau above falls that split highlands from lowlands; here the \textbf{Lence} meets the Belworth after winding west--northwest from Vhasia. Viterra also reaches the Dolmis to the east.

\subsection*{Ubral and the Dolmis Northwest}
\textbf{Ubral} occupies the northwestern Dolmis coast; passes over or under the Aelerians connect to the Mistlands. South along the Dolmis coast, Belworth-mouth shipping with Fairport/Marcott ties this corridor to Viterra/Vhasia traffic.

\subsection*{Thepyrgos and the Astroegro Straits}
\textbf{Thepyrgos} (city and province) spans a short peninsula and archipelago south of Vhasia/Viterra, controlling the Astroegro Straits and pivoting between the Dolmis (northeast) and the Amaranthine (south).

\subsection*{The Aelerian Crownlands and the Mistlands}
The \textbf{Aeler} hold the mountains north of Vhasia, Viterra, and Ubral, with gates and under-vaults that regulate access to the \textbf{Mistlands}---the protectorate that stretches from the Direwood's shadow to the Aberderrin coast (where \textbf{Payden's Port} stands).

\subsection*{Northern Coasts and Steppes}
The \textbf{Linn} hold the north coasts and raid down the Yloka into the Aberderrin and Dolmis. The Violet Steppes & Meadows (\textbf{Ykrul} and \textbf{Vilikari} frontiers) lie northwest above Ecktoria and Acasia.

\subsection*{Eastern Woods and Hills}
Across the Aberderrin from the Mistlands stands the \textbf{Valewood}. South along the Dolmis coast are the \textbf{Aelinnel} lands; further south roll the \textbf{Aelaerem} hills.

\section{Core Travel Procedure}
For each \textbf{leg} of a journey, draw 3--4 cards using the decks for your destination and controlling authority.
\begin{itemize}
  \item \SuitSpade{} from the \textbf{destination deck}: sets the scene (place).
  \item \SuitHeart{} from the \textbf{destination deck}: introduces the local actor or faction.
  \item \SuitClub{} from the \textbf{Wilds} (general hazards) or destination (if strongly policed): brings pressure.
  \item \SuitDiamond{} from the \textbf{authority that gates the route}: papers, escorts, rights, or exceptions.
\end{itemize}
Set a travel clock by the highest rank (\ClockSizes). On success, advance to the next leg; on failure, mark delay, debt, or diversion and resolve a consequence in the fiction.

\subsection*{Mode rules}
\begin{itemize}
  \item \textbf{Sea legs (Amaranthine/Dolmis/Aberderrin).} If Theona or Valewood \emph{9s} show up anywhere in the seed, add an omission or taboo to the leg (a missing step, an unsaid name). Linn \SuitClub{} draws tend to add booms, fog, or muster.
  \item \textbf{Passes & Underways (Aeler).} Any \emph{A} may convert a surface route to an under-route. \SuitDiamond{} from Aeler always codifies outcomes---no roll---for sealed doors, breath time, and escorted segments.
  \item \textbf{Rivers.} Bridges, booms, and law in Ecktoria/Viterra; reed-mazes and bell-lines in Mistlands/Linn waters.
  \item \textbf{Frontier blends.} When origin and destination disagree on law, draw two \SuitDiamond{}s (one from each law) and choose which you will be judged by at the end of the leg.
\end{itemize}

\section{Route Modules}
\subsection{Amaranthine Coastway}
Kahfagia \textrightarrow{} Ecktoria \textrightarrow{} Acasia \textrightarrow{} Marcott (Vhasia) \textrightarrow{} Fairport (Viterra).
\begin{itemize}
  \item Draw \SuitSpade{}+\SuitHeart{} from the \emph{port you are entering}.
  \item Draw \SuitClub{} from the Wilds (gales, quarantine, harbor booms).
  \item Draw \SuitDiamond{} from the \emph{port's deck} (permits, berths), or \emph{Kahfagia} if convoy/lantern law applies.
\end{itemize}
\emph{Note:} \textbf{Silkstrand} is Acasia's sole metropolis and a major trading hub; even land caravans hazard the marches to reach it.

\subsection{Astroegro Straits}
Thepyrgos controls the hinge between seas.
\begin{itemize}
  \item Draw \SuitSpade{}+\SuitHeart{} and \SuitClub{} from \emph{Thepyrgos} (stairs, boom, synod).
  \item Draw \SuitDiamond{} from \emph{Thepyrgos} (harbor pass, crane allotment).
  \item Switch to Dolmis/Amaranthine modules once the boom is lowered.
\end{itemize}

\subsection{Dolmis Circuits}
Fairport (Viterra) \textrightarrow{} Theona (Three Greens) \textrightarrow{} Ubral fjords \textrightarrow{} Aelinnel west shore.
\begin{itemize}
  \item \SuitSpade{}+\SuitHeart{} from destination; \SuitClub{} from \emph{Linn} or the Wilds; \SuitDiamond{} from \emph{Theona} when island custom matters.
\end{itemize}

\subsection{Aelerian Passes & Underways}
Vhasia/Viterra/Ubral south slopes \textrightarrow{} Aeler gates \textrightarrow{} Mistlands.
\begin{itemize}
  \item \SuitSpade{}+\SuitHeart{} and \SuitClub{} from \emph{Aeler}; \SuitDiamond{} from Aeler (Underway Pass, Key-Writ, Breath Time).
\end{itemize}

\subsection{Shadow Corridors}
\textbf{Thin Shore (Valewood east coast)}: risky misted corridor north--south toward Zakov.
\begin{itemize}
  \item \SuitSpade{} from Valewood, \SuitHeart{} from Mistlands (or Valewood), \SuitClub{} from Mistlands (wraiths, bell-line failures), \SuitDiamond{} from either Mistlands (Ward-salt, Lantern Writ) or Valewood (Way-cord, Truce-bough) depending on which law you invoke.
  \item Rule of 9s: any \emph{9} in the seed adds an omission.
\end{itemize}
\textbf{Green Gate hops (Aelinnel)}: tide-timed arches.
\begin{itemize}
  \item \SuitSpade{}+\SuitClub{} from Aelinnel (tides, geasa), \SuitDiamond{} from Aelinnel (Tide-path Key, Host-rights).
\end{itemize}

\subsection{River Roads}
\textbf{Belworth:} forms the boundary between Vhasia and Viterra. Twin ports at the mouth (Marcott and Fairport); at mid-reach \textbf{Tarlington Falls} requires lift/portage up to the plateau town. The \textbf{Lence} joins here from Vhasia. Use Viterra for east-bank audits/bridges, Vhasia for west-bank politics.
\textbf{Yloka:} raids run down from the north into the Aberderrin and Dolmis---overlay \emph{Linn} \SuitClub{}s and draw destination for \SuitSpade{}+\SuitHeart{}.
\textbf{Acasian rivers:} hazardous but profitable approaches to Silkstrand; use \emph{Acasia} \SuitClub{}s (feuds/levies/curse) and \SuitDiamond{}s (condotta, bridge rights).

\subsection{Steppe Frontiers (Violet Steppes & Meadows) (Violet Steppes & Meadows)}
Ykrul \leftrightarrow{} Vilikari \leftrightarrow{} Ecktoria/Acasia borders.
\begin{itemize}
  \item \SuitSpade{} from the road you ride (Wolf Road or Foedus Stone), \SuitHeart{} from the counterpart (envoys, comitatus), \SuitClub{} from Wilds or frontier decks (rasputitsa, hostage protocol, foedus recall), \SuitDiamond{} from either \emph{Ykrul} (paiza, remounts) or \emph{Vilikari} (foedus seal, bench-right).
\end{itemize}

\section{Gateways & Control Points}
\begin{longtable}{@{}p{0.23\linewidth}p{0.19\linewidth}p{0.22\linewidth}p{0.3\linewidth}@{}}
\toprule
\textbf{Gate} & \textbf{Deck (Spade/Heart)} & \textbf{Diamond Source} & \textbf{Typical Leverage}\\
\midrule\endhead
Qeresh-on-Sea (Kahfagia) & Kahfagia & Kahfagia & Convoy letter; lantern-law warrant; pilotage token.\\
Ecktoria Capital Port & Ecktoria & Ecktoria & Berth priority; coin-house credit; procession permit.\\
Silkstrand & Acasia & Acasia & Sole metropolis; Exchange floor; condotta; watergate priority; caravan corridor.\\
Marcott & Vhasia & Vhasia & Letters patent; safe-conduct; fairground precedence.\\
Fairport & Viterra & Viterra & Customs seal; ferry priority; Dawn escort.\\
Tarlington Falls (plateau lift) & Viterra & Viterra & Lift priority; bridge/portage pass; County Thing ruling.\\
Thepyrgos (boom) & Thepyrgos & Thepyrgos & Harbor pass; crane allotment; Synod indulgence.\\
High-Mist Pass & Aeler & Aeler & Underway pass; Key-Writ; Breath time allotment.\\
Payden's Port & Mistlands & Mistlands & Lantern writ; ward-salt; wraith indemnity.\\
Thin Shore (Valewood) & Valewood & Valewood/Mistlands & Way-cord; dew-mirror; lantern writ.\\
Green Gate (Aelinnel) & Aelinnel & Aelinnel & Tide-path key; host-right bough; salt license.\\
Ubral Passes & Ubral & Aeler/Ubral & Vurim pass-ring; hill-fort shelter writ; guide's braid.\\
Thing-holm (Linn) & Linn & Linn & Thing ruling; harbor-green; escort writ.\\
Foedus Stone (March) & Vilikari & Vilikari/Ykrul & Foedus renewal; remount allotment; hostage treaty.\\
Zakov Roadstead & (Local) & Theona/Linn & Raid-truce ribbon; whale-road escort; market immunity.\\
\bottomrule
\end{longtable}

\section{Special Rules & Taboos}
\subsection*{Theona & Valewood 9s}
Whenever a \emph{9} appears in a seed involving Theona or Valewood, add an omission: a missing step, an unsaid name, an unseen guest. If the \emph{9} is a \SuitDiamond{}, you may break the taboo once---someone will come to collect.

\subsection*{Aeler Aces}
An \emph{A} can rewire routes beneath the mountains. You may swap your \SuitSpade{} to Aeler (keeping the other seed cards) and continue under-vault procedures.

\subsection*{Echoing Omens}
Any \emph{A} adds a motif you can echo later (ash on the wind, a bell heard underground, a landmark that returns elsewhere). Make the world feel like it remembers.

\section{Worked Itineraries}
\subsection*{A) West-to-East Coastal Haul (Kahfagia \textrightarrow{} Viterra)}
\begin{Leg}{Leg 1: Qeresh-on-Sea \textrightarrow{} Ecktoria}
\textbf{Seed:} Ecktoria \SuitSpade{} (Aqueduct arcades), Ecktoria \SuitHeart{} (Coin-house factor), Wilds \SuitClub{} (gale), Kahfagia \SuitDiamond{} (convoy letter).\\
\textbf{Clock:} 6. \textbf{Read:} cross under coin-house eyes with convoy priority; bad weather strains schedules.
\end{Leg}
\begin{Leg}{Leg 2: Ecktoria \textrightarrow{} Silkstrand}
\textbf{Seed:} Acasia \SuitSpade{} (Three-Queens Bridge), Acasia \SuitHeart{} (Dyers' Guildmistress), Acasia \SuitClub{} (loom strike), Acasia \SuitDiamond{} (Exchange pass).\\
\textbf{Clock:} 6--8 depending on unrest.
\end{Leg}
\begin{Leg}{Leg 3: Silkstrand \textrightarrow{} Marcott}
\textbf{Seed:} Vhasia \SuitSpade{} (Pont-du-Tithe), Vhasia \SuitHeart{} (Parlement clerk), Vhasia \SuitClub{} (coin rumor), Vhasia \SuitDiamond{} (letters patent).
\end{Leg}
\begin{Leg}{Leg 4: Marcott \textrightarrow{} Fairport}
\textbf{Seed:} Viterra \SuitSpade{} (Fairport tideworks), Viterra \SuitHeart{} (shipwright), Linn \SuitClub{} (boom lifts), Viterra \SuitDiamond{} (customs seal).\\
\textbf{Clock:} 6.
\end{Leg}

\subsection*{B) Acasia \textrightarrow{} Mistlands (Forgotten Pass + Under-Gate)}
\begin{Leg}{Leg 1: Silkstrand \textrightarrow{} Aeler Gate}
\textbf{Seed:} Aeler \SuitSpade{} (Avalanche gallery), Aeler \SuitHeart{} (Geometer), Aeler \SuitClub{} (Engineer requisition), Aeler \SuitDiamond{} (Underway Pass).
\end{Leg}
\begin{Leg}{Leg 2: Gate \textrightarrow{} Mistlands}
\textbf{Seed:} Mistlands \SuitSpade{} (Bell-Line levee), Mistlands \SuitHeart{} (Bell-warden), Mistlands \SuitClub{} (wraith crossing), Mistlands \SuitDiamond{} (Ward-salt).
\end{Leg}

\subsection*{C) Thin Shore \textrightarrow{} Zakov Corridor}
\begin{Leg}{Leg 1: Payden's Port \textrightarrow{} Thin Shore}
\textbf{Seed:} Valewood \SuitSpade{} (Green lane), Mistlands \SuitHeart{} (Protectorate clerk), Mistlands \SuitClub{} (bell failure), Mistlands \SuitDiamond{} (Lantern writ).
\end{Leg}
\begin{Leg}{Leg 2: Thin Shore transit}
\textbf{Seed:} Valewood \SuitSpade{} (Unfound stile), Valewood \SuitHeart{} (Path-warden), Valewood \SuitClub{} (Sweet wind), Valewood \SuitDiamond{} (Way-cord).\\
\textbf{Rule of 9s} applies.
\end{Leg}

\subsection*{D) Frontier Mediation (Ykrul \leftrightarrow{} Vilikari)}
\begin{Leg}{Leg: Foedus Stone parley}
\textbf{Seed:} Vilikari \SuitSpade{} (Foedus Stone), Ykrul \SuitHeart{} (Noyan envoy), Wilds \SuitClub{} (rasputitsa), Vilikari \SuitDiamond{} (Foedus Renewal).\\
\textbf{Blend:} also draw a Ykrul \SuitDiamond{} (Paiza). Choose which law saves your schedule.
\end{Leg}

\section{Pace, Clocks, and Consequences}
\begin{itemize}
  \item \textbf{Road/river leg:} 6-segment clock; +2 segments if crossing a law boundary (new passports, new prayers).
  \item \textbf{Sea strait/mountain pass:} 8-segment clock; --2 segments if you spend a \SuitDiamond{} that codifies priority (pilotage, pass, convoy).
  \item \textbf{Shadow corridor:} 6-segment clock; on any \emph{A}, add a free omen to echo later.
  \item \textbf{Failure defaults:} delay (lose time; an enemy advances), diversion (alternate route; new \SuitClub{}), or debt (promise, fine, or favor owed to the authority that controls the gate).
\end{itemize}

\section{Appendix A: Nomenclature (Quick Canon)}
\begin{multicols}{2}
\begin{itemize}
  \item \textbf{Kahfagia} — western convoys & lantern law.
  \item \textbf{Ecktoria} — central peninsula; coin-houses & forums.
  \item \textbf{Acasia} — broken marches; \textbf{Silkstrand} the sole metropolis and Amaranthine hub.
  \item \textbf{Vhasia} — fractured sun; Marcott at the Belworth mouth.
  \item \textbf{Viterra} — last kingdom; Fairport at the Belworth mouth; Tarlington on the falls.
  \item \textbf{Ubral} — stone between spears; passes to the Dolmis NW.
  \item \textbf{Thepyrgos} — stairs, synod, and the Astroegro Straits.
  \item \textbf{Aeler} — crowns & under-vaults across the Aelerians.
  \item \textbf{Mistlands} — bells, salt, and breath; Payden's Port.
  \item \textbf{Valewood} — empire under leaves; thin shore.
  \item \textbf{Aelinnel} — stone, bough, and bright tide-gates.
  \item \textbf{Aelaerem} — hearth & hollow across rolling downs.
  \item \textbf{Linn} — skerries & storm-oaths; the Yloka road.
  \item \textbf{Ykrul} — wolf standards & winter camps.
  \item \textbf{Vilikari} — laurels & longhouses.
  \item \textbf{Amaranthine Sea} — western sea.
  \item \textbf{Dolmis Sea} — inner sea to the east.
  \item \textbf{Aberderrin Sea} — northeastern basin.
  \item \textbf{Astroegro Straits} — hinge between seas.
\end{itemize}
\end{multicols}

\bigskip
\noindent\textit{Use this as a GM screen: deal the leg, say the place, introduce the actor, let the pressure bite, and hold the leverage card visible. When a \SuitDiamond{} is played, the world moves.}

\end{document}

