\documentclass[11pt,letterpaper,twoside]{book}

% =========================
% PACKAGES
% =========================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
≥ometry{
  letterpaper,
  inner=1.5in,
  outer=1in,
  top=1in,
  bottom=1in,
  bindingoffset=0.5cm
}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\tcbuselibrary{skins,breakable,listings}
\usepackage{makeidx}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{etoolbox}
\usepackage{graphicx}
\usepackage{float}
\usepackage{calc}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage{chngcntr}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{tabularray}

% =========================
% COLOR DEFINITIONS
% =========================
\definecolor{FePrimary}{HTML}{3F6E8C}
\definecolor{FeSecondary}{HTML}{2A4D6A}
\definecolor{FeAccent}{HTML}{C44536}
\definecolor{FeLightBg}{HTML}{F5F7FA}
\definecolor{FeDarkText}{HTML}{1A2F3E}
\definecolor{FeMidGray}{HTML}{6C7A89}
\definecolor{FeRow}{HTML}{F5F7FA}

% =========================
% HYPERREF SETUP
% =========================
ℎypersetup{
  colorlinks=true,
  linkcolor=FePrimary,
  citecolor=FePrimary,
  urlcolor=FeAccent,
  pdftitle={Fate's Edge: Comprehensive Resource Guide},
  pdfauthor={Nicholas A. Gasper},
  pdfsubject={TTRPG Reference Document},
  pdfkeywords={Fate's Edge, TTRPG, reference, rules}
}

% =========================
% CUSTOM COMMANDS
% =========================
≠wcommand{\FEsc}[1]{\textsc{#1}}
≠wcommand{\fatebox}[2][colback=FeLightBg, colframe=FePrimary]{
  \begin{tcolorbox}[#1, sharp corners, boxrule=0.5pt, arc=0pt, left=6pt, right=6pt, top=6pt, bottom=6pt]
    #2
  \end{tcolorbox}
}
≠wcommand{\feTableStart}{\rowcolors{2}{FeRow}{white}\small}
≠wcommand{\feTableEnd}{\rowcolors{2}{}{}}

% Title page commands
≠wcommand{\feSetTitle}[1]{\def\fetitle{#1}}
≠wcommand{\feSetSubtitle}[1]{\def\fesubtitle{#1}}
≠wcommand{\feSetTagline}[1]{\def\fetagline{#1}}
≠wcommand{\feSetVersion}[1]{\def\feversion{#1}}
≠wcommand{\feSetPublisher}[1]{\def\fepublisher{#1}}
≠wcommand{\feSetCoverArt}[1]{\def\fecoverart{#1}}

≠wcommand{\feTitlePage}{%
  \begin{titlepage}
    ¢ering
    \vspace*{2cm}
    {\Huge\bfseries\sffamily \fetitle \par}
    \vspace{1cm}
    {Łargeℹtshape \fesubtitle \par}
    \vspace{1.5cm}
    {łarge \fetagline \par}
    \vfill
    {łarge Version \feversion \par}
    \vspace{0.5cm}
    {\normalsize \fepublisher \par}
    ℹfdefined\fecoverart
      \vspace{1cm}
      ∈cludegraphics[width=0.6\textwidth]{\fecoverart}
    \fi
    \vspace{1cm}
  \end{titlepage}
}

% Index and cross−reference setup
\makeindex
\counterwithin*{section}{chapter}
\counterwithin*{subsection}{section}

% =========================
% HEADERS & FOOTERS
% =========================
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\slshape ≤ftmark}
\fancyhead[RE,LO]{\slshape \rightmark}
\fancyfoot[CE,CO]{þepage}
\fancyfoot[LE,RO]{\footnotesize \textit{Fate's Edge: Comprehensive Resource Guide}}
\renewcommand{ℎeadrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% Chapter style
\chapterfont{\Huge\bfseries\sffamily}
\sectionfont{Łarge\bfseries\sffamily}
\subsectionfont{\normalsize\bfseries\sffamily}

% =========================
% BEGIN DOCUMENT
% =========================
\begin{document}

% Title info
\feSetTitle{Fate's Edge: Comprehensive Resource Guide}
\feSetSubtitle{Comprehensive Reference Document}
\feSetTagline{Risk, consequence, and sharp choices at the table.}
\feSetVersion{v0.9}
\feSetPublisher{}
% \feSetCoverArt{art/cover.jpg} % optional
\feTitlePage

% Front matter
\frontmatter
\tableofcontents

% Main matter
\mainmatter
∈clude{reviewer_guidance}
≠wpage
∈clude{sections/01_coreₚrinciples}
≠wpage
∈clude{sections/02_coreₘechanics}
≠wpage
∈clude{sections/03_deck_generators}
≠wpage
%∈clude{sections/04ₜravelᵣeference}
%≠wpage
∈clude{sections/05ᵣesourceₘanagement}
≠wpage
∈clude{sections/06_characterₐdvancement}
≠wpage
∈clude{sections/07ₘagicₛystem}
≠wpage
∈clude{sections/08_combat_conflict}
≠wpage
∈clude{sections/09ₗoreₕeavy}
≠wpage
∈clude{sections/10ₜoolsₜips}
≠wpage
∈clude{sections/11ₐppendices}
≠wpage

\chapter*{Closing Thoughts}
\addcontentsline{toc}{chapter}{Closing Thoughts}
∈dex{Closing Thoughts}

Every roleplaying game is, at its heart, a shared act of imagination. \FEsc{Fate's Edge} offers rules and structures, but the real magic happens between players and game master, in the pauses between dice rolls and the sparks that leap across the table. The mechanics are not a cage−−−they are a scaffold, meant to support the story you want to tell.

You will not find every answer written in these pages, nor should you. The best sessions arise when the table embraces uncertainty: when a Story Beat pulls the tale in an unexpected direction, when a character's flaw drives the narrative forward, when laughter follows disaster. The system gives you tools; the table gives those tools meaning.

If you remember nothing else, carry these three principles into play:
\begin{itemize}
  ℹtem \textbf{Fiction First.} Let the world breathe, and let the rules bend to the needs of the story.
  ℹtem \textbf{Every Choice Matters.} Risk, consequence, and reward shape characters as surely as dice.
  ℹtem \textbf{Build Together.} Collaboration turns scenes into sagas and mechanics into myth.
\end{itemize}

The rules are here to help you weave drama, not to constrain it. Use them generously, ignore them when they slow the story, and trust your table to improvise with honesty and care. Above all, let the edges of fate guide you not toward certainty, but toward possibility.

\vfill
\begin{center}
ℹtshape
May your journeys be bold, your dice fall true, and your stories echo long after the table is cleared.
\end{center}

% Back matter
\backmatter
\printindex

\end{document}
