\section*{Character Profile: Kes Orren, "The Tollkeeper"}

\textbf{Alias:} The Tollkeeper, Kes "Underhill" Orren, Merchant Factor K. Orendale \\
\textbf{Affiliation:} Independent Operator (Primary); rumored ties to the Reed-House Vilikari \\
\textbf{Race:} Half-Elf (Vilikari / Lethai-ar descent) \\
\textbf{Age:} 38 \\
\textbf{Appearance:} Purposefully unremarkable: medium height, lean build. Mousy brown hair worn to hide subtly pointed ears beneath a practical wool cap. Weather-stained merchant’s garb in muted tones, a battered leather satchel, and gloves worn even indoors. Up close, a quiet Lethai cast to the eyes and a faint river-freckle across the nose. A small scar marks her left hand from a “misplaced ledger.” \\
\textbf{Demeanor:} Calm, soft-spoken, meticulous. Projects the air of a nervous, overworked clerk. Speaks in precise, pedantic terms; eye contact never lingers. Habitually adjusts her gloves. \\
\textbf{Mannerisms:} Constantly checks a small, ornate (non-functional) pocket watch. Writes in a small notebook she carries everywhere. Counts on her fingers when thinking, often under the table.

\subsection*{Background}

Kes Orren was not born to smuggling. She was the third daughter of a minor Vilikari factor in the river towns, trained in ledgers, tariffs, and the sacred \emph{Two Ledgers, One Truth}. Her early years passed in counting houses along the Ecktorian frontier, keeping goods, coin, and favors flowing—smoothly, and above all, \emph{profitably}.

Kes possessed an uncanny sense for \emph{gaps}—not only in books, but in laws, routes, and people. “Creative” tariff work for clients became, by degrees, the discreet movement of goods (and sometimes people) that empires and hosts would rather not see cross a border. She learned to be \emph{forgettable} and to wield a bland, honest face like armor.

Leaving her family house, Kes built an independent network on punctuality, procedure, and price. Her reputation rests on two pillars: she is never late, and her routes are costed to the breath. The moniker “Tollkeeper” stuck not because she \emph{collects} tolls, but because she can \emph{price} any crossing—legal or otherwise—before the lamp burns down.

\subsection*{Core Concept}

Kes embodies Vilikari pragmatism at the margins. No dashing smuggler or violent fixer, she is a master of \emph{logistics, timing, and legal sleight-of-hand}. Frontier regulations, seasonal rights, and faction tensions form a maze only she seems to navigate without incident. Her power is making the impossible look routine.

\subsection*{Key Relationships}

\begin{itemize}
    \item \textbf{Ykrul Judges (Border Ford):} Formal, respectful rapport. Dues paid punctually; Kes has drafted seasonal concessions benefiting both sides. The Judges appreciate her reliability and Bowl \& Board discipline.
    \item \textbf{Black Banner Quartermasters:} Several QMs list her as a “preferred supplier.” She sources oddments, smooths personnel passage, and always knows where a unit’s line is thin (for legitimate trade, of course). A few owe small favors.
    \item \textbf{Ecktoria Customs (Varied):} Some clerks view Kes as a useful facilitator; others as a thorn. She maintains exact notes on which is which and who drinks which tea.
    \item \textbf{Ostrikari Envoys (Seasonal):} At winter moots, Kes serves as neutral broker between clans and Ecktorian merchants, honoring guest-right scrupulously and translating between “ledger law” and “host law.”
    \item \textbf{Local Aeler Stonewrights:} Quiet accounts for tool-work and wagon repairs; mutual respect for fair dealing and craft.
\end{itemize}

\subsection*{Motivations}

\begin{itemize}
    \item \textbf{Profit (Primary):} A comfortable life—on her own terms.
    \item \textbf{The Puzzle:} Routes, ledgers, and timing are games she intends to win.
    \item \textbf{Network Security:} Trust and function of her web come before pride.
    \item \textbf{Quiet Pride:} Proving a “humble” factor can outmaneuver empires through patience and wit.
\end{itemize}

\subsection*{Fate's Edge Stats}

\textbf{Attributes:} \\
Body: 2 \quad Wits: 4 \quad Spirit: 3 \quad Presence: 2

\textbf{Skills:} \\
Lore 3 \quad Subterfuge 4 \quad Insight 3 \quad Tinker 2 \quad Sway 2 \\
Notice 3 \quad Survival 1 \quad Command 1 \quad Arcana 0 \\
Athletics 1 \quad Stealth 2 \quad Ranged 1 \quad Melee 1 \quad Craft 2

\textbf{Key Talents:}
\begin{itemize}
    \item \textbf{Caster's Gift (2 XP):} Subtle utility weaves (minor illusion, chill, draft) for misdirection: blank a line of text under lamp-light, raise a fog-bead over a ledger, cool a brand to delay a seal.
    \item \textbf{Countkeeper's Eye (Custom / 4 XP):} +1 die on rolls parsing ledgers, tariffs, contracts, or hidden clauses; spot a “Two Ledgers” mismatch at a glance.
    \item \textbf{Night Roads Expert (Custom / 3 XP):} Once per leg, reroute around a major Front (checkpoint closure, patrol surge) via timing window or lesser-known path; mark \textbf{Night Road Heat +1}.
    \item \textbf{Reputation as Tool (Narrative):} Begin with \textbf{Banked Fear (1)} among contraband circles. Spend to \emph{Parley First} in smuggling disputes or to flip one enemy action to \emph{Controlled}.
\end{itemize}

\textbf{Affinity:} Vilikari / Lethai-ar (half-elf) \\
\textbf{Gift:} Gift of the Mind (logistics and legal cunning)

\textbf{Strings:}
\begin{itemize}
    \item \textbf{Reed Chit:} Stamped lock-pass for a specific checkpoint; \textbf{DV --1} on crossings/inspections there.
    \item \textbf{Toll Braid:} Last season’s dues paid; \textbf{Position +1} at toll houses/checkpoints.
    \item \textbf{Guild Widow’s Trust:} A fence will hold goods, no questions; on use, mark \textbf{City Heat +1}.
    \item \textbf{Ford Right (Season):} Bowl \& Board concession; treat a specific ford/bridge as friendly terrain.
    \item \textbf{Quartermaster’s Voucher:} IOU from a Banner QM: call 1 squad for 1 scene \emph{or} cancel one squad pursuit.
    \item \textbf{Ykrul Windbreak Right:} Seasonal right to rest/pass within a named ring-camp.
\end{itemize}

\textbf{Debts:}
\begin{itemize}
    \item \textbf{Ecktoria Audit Marker [2]:} A minor discrepancy awaits discovery if an audit deepens.
    \item \textbf{Banner QM Favor [1]:} A “small favor” promised regarding a future cargo run.
\end{itemize}

\textbf{Mechanical Hooks:}
\begin{itemize}
    \item \textbf{Lamp-Lit Paper:} Present any writ under a lit lamp and read it aloud: \textbf{Position +1} in bureaucratic scenes; risk \textbf{Night Road Heat [1]} if forged or tense.
    \item \textbf{Two Weights, One Truth:} Lay out public and private prices together: \textbf{DV --1} to close a deal; risk \textbf{Night Road Heat [1]} if concealed.
    \item \textbf{Guest Cup on the Threshold:} Pour for the guard before yourself: the first social \textbf{SB} against you may be converted to \textbf{Heat +1} instead of a complication.
\end{itemize}

\textbf{Clocks:}
\begin{itemize}
    \item \textbf{Night Road Heat [4/6]:} Customs, syndicates, rivals. On fill: raids, curfews, embargoes.
    \item \textbf{Imperial Audit [2/4]:} On fill: seizures, fines, flipped informants.
    \item \textbf{Banner Debt [1/4]:} On fill: a “discipline visit.”
    \item \textbf{Ford War [3/6]:} Competing rights; price swings with season and tension.
\end{itemize}

\subsection*{Sample Adventure Hooks}

\begin{enumerate}
    \item \textbf{The Misplaced Tithe:} A tithe bound for an Ostrikari winter moot vanished on Kes’s route. Prove theft, recover the coin, or broker peace before \emph{Fury} overrules \emph{Measure}.
    \item \textbf{The Auditor’s Ledger:} A feared customs auditor is found dead; their private ledger is missing. Kes is a prime suspect. Retrieve, erase, or redirect the “second weight” before the inquest.
    \item \textbf{The Banner’s Black Chest:} A Banner pay-chest vanished at a “safe” ford. Judges are involved; tensions rise. Find the chest, fix the blame, or write a charter all can sign without losing face (or coin).
    \item \textbf{The Silent Ferry:} A key night ferry goes dark. Syndic tight-lipped, strange lights on the far bank. Reopen the route, unlock an alternative, or convince a client to wait—and pay for the risk.
\end{enumerate}

\end{document}
