\Chapter{Legends}

\subsection{Saikou \& Amari (Corrected Build)}

\subsubsection{Saikou Ira — “The Red Lamps See All”}
\textbf{XP Path:} Mastermind \quad
\textbf{Core:} A cursed strategist who reads positions—legal, social, criminal—the way others read faces.

\paragraph{Attributes \& Core Skills}
Wits 4, Presence 3, Spirit 3, Body 2.\\
Key Skills: \textbf{Canray (Royal Game) 4}, Insight 4, Arcana 3, Command 3, Sway 3, Lore (World’s Secrets) 3.

\paragraph{What Canray Is (lore, not a combat style)}
\emph{Canray} (formal: Canré) is the royal board-game and its theory. In play, it’s your \emph{reading discipline}: modeling incentives, exits, and fake “obvious” moves. Use it to frame \emph{plans and predictions}, never to strike.

\paragraph{Talent: Fogbound Read (Canray School)}
Once/scene, after you \emph{sell} a Canray analysis (MRS: Method–Risk–Signal) of the current situation, roll \textbf{Wits + Canray}.\\
\emph{On success} choose two (on crit, choose all three):
\begin{itemize}
  \item \textbf{Exit Certainty:} ask the GM which single “safe” exit your foes think they have (truthful).
  \item \textbf{Bait Map:} declare one “obvious” reply your move appears to invite; it is a \emph{trap} (DV $-1$ vs that reply for your side this beat).
  \item \textbf{Noise Screen:} set Position to \emph{Controlled} for the next allied approach that exploits your read.
\end{itemize}
\emph{On failure:} GM banks +2 CP (\emph{you are being read}) and may raise DV +1 on the next opposed social/tactical action targeting your plan.

\paragraph{Talent: Oligarch’s Cunning (unchanged)}
Once/session, instantly name a principal leverage (want/fear) in a negotiation; gain \emph{Controlled} for the ensuing parley.

\paragraph{Prestige: The Red Lamps See All (optional swap-in)}
Once/session, under literal lamplight or equivalent \emph{tell}, ask one \emph{hidden} question about a person/place/object in view (deepest lie, true aim, or concealed channel). Answer is true but invites a personal, supernatural complication: GM banks +2 CP; your next major complication trends uncanny.

\paragraph{Asset: Kahfagian Bloodline (Major)}
A standing social engine (charters, patronage, enemies).
\begin{itemize}
  \item \textbf{Invoke:} mark 1 \emph{Exposure} (Quiet$\to$Rumored$\to$Burned) \emph{or} spend Favor/coin to pull strings (counts as a small Diamond in ports/convoys).
  \item \textbf{Backfire:} on any GM CP spend in scenes you invoked it, the matriarch’s \emph{price} arrives (debt, audit, rival).
\end{itemize}

\paragraph{Follower: Amari (Cap 4) — see below}

\subsubsection{Amari — The Loyal Sergeant}
\textbf{XP Path:} Solo \quad
Body 4, Spirit 3, Wits 2, Presence 2.\\
Key Skills: Melee 4, Athletics 3, Intimidation 3, Survival 2, Tactics 2.

\paragraph{Talents}
\textbf{Unstoppable Advance (8 XP):} When moving directly to remove a threat to Saikou, +1 effect and immunity to intimidation.\\
\textbf{Interpose (5 XP):} 1/scene, step in and take Harm meant for Saikou; roll your pool to reduce/negate.\\
\textbf{There Is No Puzzle, Only The Wall (4 XP):} Use Body+Melee to break non-combat obstacles (loudly).

\subsubsection{Table Use (tight loop)}
\begin{itemize}
  \item Saikou frames the scene with \textbf{Intricate} Canray analysis (MRS) $\Rightarrow$ \emph{Fogbound Read} roll $\Rightarrow$ sets Position/DV tweaks and truth.
  \item Amari cashes the window with direct action; if plans wobble, he \emph{Interposes} or \emph{smashes through}.
  \item \textbf{GM CP spends} to hit the bond: tempt Amari with “clean” solutions that wreck stealth; tempt Saikou with Bloodline favors that cost Exposure.
\end{itemize}

\subsubsection{GM Levers (CP menu, tailored)}
\begin{itemize}
  \item \textbf{1 CP:} A rival “reads the read”: +1 DV vs the declared bait this beat.
  \item \textbf{2 CP:} Bloodline push—gain immediate access, mark \emph{Exposure} on the Asset.
  \item \textbf{3 CP:} Red Lamp echo—convert the next mundane complication into an uncanny one bound to Saikou’s insight.
  \item \textbf{4 CP:} Mother’s writ closes a door you just opened; route changes or pay a Diamond-class concession.
\end{itemize}

\subsection{Riarhin — “Remorse in Red”}

\subsubsection{Core Read}
\textbf{XP Path:} Paragon \quad
\textbf{Core:} A high-born Lethai-thora heir whose \emph{glamour} and memory-rites shape the field—yet all gifts carry cost. Her red hair and name are prophecy-bound, and the sword \emph{Remorse} is both legacy and burden.

\paragraph{Attributes \& Core Skills}
Presence 3, Spirit 4, Wits 3, Body 2.\\
Key Skills: \textbf{Glamour (Illusion-Weave) 4}, Lore (Elven Memory) 3, Rituals 3, Sway 3, Melee (Blade) 2, Survival (Green Ways) 2.

\paragraph{What Glamour Is (lore, not battle magic)}
\emph{Glamour} is the art of seeming and sympathy—tilting attention, polishing truths, stitching mood to light. It persuades the world instead of wrenching it. It is safer than raw sorcery, but never free of price.

\subsubsection{Talents}
\paragraph{Veil-Thread (Glamour School) \normalfont(6 XP)}
Once/scene, after you \emph{sell} a glamour (tell what the senses get, what they \emph{miss}, and whose will is anchored), roll \textbf{Presence + Glamour}.\\
\emph{On success} choose two; \emph{on crit}, choose all three:
\begin{itemize}
  \item \textbf{Soft Focus:} downgrade one hostile’s Position against you (Desperate$\rightarrow$Risky, Risky$\rightarrow$Controlled).
  \item \textbf{Borrowed Light:} your next social opener gains +1 effect if it leans into the mood you set.
  \item \textbf{Kind Cut:} convert up to 1 Harm from your illusion’s backlash into +1 Fatigue instead (you carry the price).
\end{itemize}
\emph{On failure:} GM banks +2 CP; someone important notices the seam in your seeming.

\paragraph{Name-Bound (High Bough Oath) \normalfont(5 XP)}
When you invoke a sworn name (oath, lineage, treaty) in good faith, gain \emph{Controlled} for the opening exchange and ask one narrow truth about the oath’s current standing (GM answers plainly). Breaking or gaming the oath this scene gives the GM +2 CP immediately.

\paragraph{Remorse, the Balancing Edge \normalfont(8 XP)}
When you draw \emph{Remorse} and take a principled stance aloud (mercy, restraint, or rightful redress), you may treat a duel, parley, or standoff as \emph{contested but bounded}: neither side can escalate to lethal force this beat unless someone spends a \emph{Diamond} or the GM spends 3 CP to shatter the balance. If you violate your own stance this scene, you gain +2 Fatigue at scene end.

\subsubsection{Asset}
\paragraph{Major Asset: \emph{Remorse} (Relic Blade)}
A storied heirloom that \emph{answers} to duty and loss.\\
\textbf{Invoke:} mark 1 \emph{Exposure} on your legend \emph{or} accept +1 Fatigue to assert dueling terms, truce-lines, or guest-rights for one beat (GM adjudicates scope).\\
\textbf{Backfire:} when artifacts of fate or old empires are near, the blade \emph{pulls}—the GM may spend 2 CP to steer events toward that gravity.

\subsubsection{Prestige (Optional)}
\paragraph{Mistweft Heir (Prestige, 15+ XP)}
Once per session, you may ask the world a \emph{why} at a site of deep memory (ruin, barrow, oath-stone). Receive a true answer framed as a remembered scene or proverb; immediately bank +1 Fatigue and the GM banks +2 CP (\emph{memory answers, but it takes}).\\
\textbf{Epic Hook — The Burden of Remorse:} when you call on Remorse to end a larger conflict cleanly (siege, riot, blood-feud), the GM may dump any CP to \emph{balance the ledger elsewhere}: the strife dissolves here, but its unpaid cost migrates to a person or place you care about.

\subsubsection{Table Use (tight loop)}
\begin{itemize}
  \item \textbf{Open:} Pitch an \textbf{Intricate} glamour (tell the seam) $\Rightarrow$ \emph{Veil-Thread} to shape Position/effect.
  \item \textbf{Press:} Name an oath or lineage for \emph{Name-Bound} insight/control.
  \item \textbf{Close:} Draw \emph{Remorse} to set dueling bounds or truce-lines; spend \emph{Diamonds} to hold them under fire.
\end{itemize}

\subsubsection{Scene Keys \& GM Levers}
\paragraph{Keys}
Ruin with Living Memory; Court beneath the Bough; Treaty-Stone by the Ford; Barrow door at noon.
\paragraph{CP Menu (examples)}
\begin{itemize}
  \item \textbf{1 CP:} The glamour leaves a tell—someone fixates on the wrong detail.
  \item \textbf{2 CP:} \emph{Remorse pulls}: a relic’s presence bends the route; add a detour clock.
  \item \textbf{3 CP:} An elder calls a prior oath that narrows your options now.
  \item \textbf{4+ CP (Epic Hook):} \emph{Balance Elsewhere}—today’s peace seeds tomorrow’s grief where you cannot stand guard.
\end{itemize}
