\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{mdframed}

\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}

\newenvironment{talentbox}[1]{%
  \begin{mdframed}[backgroundcolor=gray!10, linewidth=1pt, linecolor=gray!50]
  \textbf{#1}\\
}{\end{mdframed}}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Fate's Edge - Noetic Knight},
    pdfauthor={Playtest Design}
}

\begin{document}

\title{Fate's Edge --- Noetic Knight (Psionic Approach)}
\author{}
\date{}
\maketitle

\section*{Core Concept}

A Noetic Knight channels disciplined thought into tangible force: shaping a mind-blade, reading intent along the thoughtweft, and bending vectors with precise telekinesis. Built on the Psionics expansion with small, themed tweaks.

\section*{Starting Build (30 XP)}

\textbf{Attributes:}
\begin{itemize}[leftmargin=*]
    \item Body 2, Wits 3, Spirit 4, Presence 2
\end{itemize}

\textbf{Skills:}
\begin{itemize}[leftmargin=*]
    \item Melee 1, Athletics 2, Insight 2, Survival 1
\end{itemize}

\textbf{Talents:}
\begin{itemize}[leftmargin=*]
    \item Psionics (4 XP) --- Access to psionic arts (skill level 2)
    \item Telekinesis (Art) --- Move/shape matter at range
    \item Telepathy (Art) --- Read surface thoughts, send impressions
    \item Foresight (Art) --- Glimpses of immediate danger
    \item Aegis of Will (Art) --- Resist/deflect mental intrusions
    \item Mental Fortress (4 XP) --- +1 die to resist psychic effects; convert psychic Harm $\rightarrow$ Fatigue
    \item Thought Thief (4 XP) --- Refined telepathy for motives, tells, and veiled intent
\end{itemize}

\textbf{Remaining XP:} 10 XP (bank for play)

\section*{Core Mechanics}

\subsection*{Combat Approach}
\begin{itemize}[leftmargin=*]
    \item Mind-Blade Manifestation: Shape a psionic edge of focused intent.
    \item Position: Often Controlled/Standard; Dominant is achievable with timely Foresight.
    \item Effect: Standard on basic strikes; Great with kinetic flourishes or setup advantages.
\end{itemize}

\subsection*{Psionic Techniques}

\begin{talentbox}{Mind-Blade (DV 3--4)}
    \begin{itemize}[leftmargin=*]
        \item Manifest a weapon of condensed will.
        \item Harm 2 (Melee), tags: [NOETIC], [FINESSE].
        \item Throw at Near range as a shimmering javelin of thought.
        \item Major use: on Partial, GM gains +1 SB; on Miss, +2 SB.
    \end{itemize}
\end{talentbox}

\begin{talentbox}{Kinetic Wave (DV 3--4)}
    \begin{itemize}[leftmargin=*]
        \item Short, directed burst to shove/pull, scatter ranks, or yank objects.
        \item Increase DV for area effects or precise disentangling.
        \item Defensive application: open space, disengage, or knock missiles aside.
        \item Major use: Partial +1 SB, Miss +2 SB.
    \end{itemize}
\end{talentbox}

\begin{talentbox}{Foresight (DV 3)}
    \begin{itemize}[leftmargin=*]
        \item Read the near-future vector of violence.
        \item +1 Position on defense for the scene; brief narrative tell of incoming threats.
        \item Consequence: mark 2 Fatigue.
    \end{itemize}
\end{talentbox}

\begin{talentbox}{Thoughtweft Read (DV 4--5)}
    \begin{itemize}[leftmargin=*]
        \item Skim surface thoughts and emotional spikes; detect deception/hostility.
        \item DV scales with target's resolve, training, or counter-wards.
        \item Consequence: mark 2 Fatigue; on Miss, GM gains +1 SB.
    \end{itemize}
\end{talentbox}

\begin{talentbox}{Aegis of Will (reaction)}
    \begin{itemize}[leftmargin=*]
        \item Interpose telekinetic filaments to spoil trajectories.
        \item +1 die to defend vs ranged attacks; convert incoming Ranged Harm 1 $\rightarrow$ Fatigue.
    \end{itemize}
\end{talentbox}

\section*{Special Rules}

\begin{talentbox}{Noetic Sense (Free, 1/scene)}
    Spend 1 Boon to step Position up by one on Perception/Insight. Can detect hidden presences, danger hum, or strong emotional residue.
\end{talentbox}

\begin{talentbox}{Trance Recovery}
    In downtime, clear +1 Mental Fatigue via disciplined meditation. Regain 1 Mental Fatigue per rest period.
\end{talentbox}

\begin{talentbox}{Overchannel (1/scene)}
    Spend 3 Fatigue for +1 Effect on a psionic action (attack or defense).
\end{talentbox}

\begin{talentbox}{Threshold Gambit}
    At 4+ Fatigue, you may push Effect +1; mark 2 Harm and generate 2 SB.
\end{talentbox}

\section*{Mental Fatigue Track}
\begin{itemize}[leftmargin=*]
    \item Mental Fatigue = 2 $\times$ Spirit = 8 segments.
    \item Mark for DV 4+ psionics, sustained multi-target control, or prolonged exertion.
    \item At 2 Fatigue: $-1$ die to psionic rolls.
    \item At 4 Fatigue: choose Overchannel, Economize (treat next DV as $-1$), or Reset (drop Position by one to clear 1 Fatigue).
    \item Overflow: when full, increase current Harm one level, then clear all Fatigue.
\end{itemize}

\section*{Advanced Talents (Progression)}

\begin{talentbox}{Kinetic Savant (7 XP)}
    +1 die on Telekinesis attacks; lift one weight class heavier without DV increase.
\end{talentbox}

\begin{talentbox}{Astral Strider (8 XP)}
    Projected awareness requires no concentration; move at normal pace in astral scouting.
\end{talentbox}

\begin{talentbox}{Mentiphage (8 XP)}
    When dealing psychic Harm, heal 1 Fatigue; 1/scene drain 1 Fatigue from a target on telepathic contact.
\end{talentbox}

\begin{talentbox}{Chorus of Thought (12 XP)}
    +2 dice to telepathic effects; broadcast calm, commands, or imagery to your whole crew.
\end{talentbox}

\begin{talentbox}{Reality Threading (15 XP)}
    1/session, bend a local causal thread (brief terrain tilt, improbable opening). Cost: mark 2 Harm and 2 Fatigue; GM frames side-effects.
\end{talentbox}

\section*{Equipment}

\begin{talentbox}{Noetic Prism (Focus)}
    \begin{itemize}[leftmargin=*]
        \item Tags: [CONDUIT], [AMPLIFY]
        \item Effect: Reduce Telekinesis Fatigue cost by 1 (min 1).
        \item Special: Sings when future vectors converge; aids Foresight narration.
    \end{itemize}
\end{talentbox}

\begin{talentbox}{Contemplative Mantle (Robes)}
    \begin{itemize}[leftmargin=*]
        \item Light armor; Harm 1 $\rightarrow$ Fatigue conversion once/scene.
        \item Tags: [MEDITATIVE], [COMFORT]
        \item Special: Advantage on Trance Recovery in quiet venues.
    \end{itemize}
\end{talentbox}

\section*{Sample Combat Line}

Opening: Controlled/Standard with Mind-Blade.

Strong Hit: Great Effect; immediately Kinetic Wave to reposition.

Partial: Success; mark 1 Fatigue, GM +1 SB.

Miss: No progress; mark 2 Fatigue, GM +2 SB.

\subsection*{Defensive Rhythm:}
\begin{itemize}[leftmargin=*]
    \item Pre-empt with Foresight to lift Position.
    \item Use Aegis of Will vs volleys; convert chip Harm $\rightarrow$ Fatigue.
    \item Raise Aegis or Mind Shield when facing hostile telepaths.
\end{itemize}

\section*{Strain \& Backlash}

\subsection*{Noetic Strain}
Fatigue is the thermal cost of cognition-as-force. High Fatigue blunts clarity and invites mistakes.

\subsection*{Cognitive Backlash}
All 1s on psionic rolls feed GM SB; overreach causes real Harm and narrative complications (echoes, migraines, sensory bleed).

\section*{Notes for GMs}
\begin{itemize}[leftmargin=*]
    \item Keep Position/DV honest: the knight excels when they earn advantage through scouting, set-up, and Foresight, not permanent Dominant stance.
    \item Let SB from misses fuel environmental shifts (loose scaffolds, flickering lanterns, panic in crowds) that the Noetic Knight can answer creatively with Telekinesis/Aegis instead of just trading Harm.
\end{itemize}

\end{document}
