```latex
\begin{followerbox}{The Wandering Ronin — \emph{Kaito the Bladeless}}
Cap: 4 \quad Scale: I \\
Playbook: Combat Ally \\
Capabilities: Iaijutsu mastery, philosophical dueling, silent movement \\
Edges: Perfect draw, read intent, blade-work sans blade \\
Liabilities: Honor code refuses unfair fights, speaks in riddles \\
Strings: Shrine blessings; duel-sanction rights; master's unfinished letter \\
Costs: Oath of the Way; no permanent service; must teach one student \\
Tracks: Loyalty 5/6, Morale 4/6, Condition 0/3 \\
Tags: Elite Training, Honor-Bound, Silent Step, Sword Saint

\paragraph{Philosophy} 
"I carry no blade, yet all swords answer to me. In stillness, I cut. In silence, I speak. The duel is a poem—let us write the final verse together."

\paragraph{Orders}
\textbf{The Unsheathed Moment:} Perfect iaijutsu strike; on hit, opponent suffers \textbf{Effect +1} to all actions against them (they move first, poorly). On strong hit, also \emph{Disarmed [1]} or \emph{Shamed [2]}. \\
\textbf{Way of No Sword:} Teach or test; on hit, grant ally \textbf{Position +1} on next social action OR impose \emph{Hesitation [2]} on one foe who witnessed the lesson. \\
\textbf{Silent Passage:} Move unseen through hostile ground; on partial, leave a poem that reveals your path (Position +1 to pursuers).

\paragraph{Special Rules}
\textbf{Bladeless Technique:} Cannot be disarmed; ignores armor against precision strikes. \\
\textbf{Master's Debt:} Once per campaign, must complete a duel tied to their master's unfinished quest. \\
\textbf{Student's Price:} If reduced to \textbf{Loyalty 0}, departs to teach rather than betray—but grants the party one major clue/boon first.

\paragraph{Integration Notes}
Perfect for: Allegiances events (refuses dishonorable patrons), Honor/Pragmatism dial (pulls party toward Honor), high-stakes duels, philosophical moments. Use in scenes requiring precision strikes, social tension, or moral guidance. Their presence improves Position +1 in any formal duel or honor-bound situation.
\end{followerbox}
```
