% !TEX program = pdflatex
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=black}

\setlist{noitemsep,topsep=3pt,itemsep=2pt,parsep=0pt}
\titleformat{\section}{\large\bfseries}{}{0pt}{}

\begin{document}

% -------------------------------------------------------------
% HEADER
% -------------------------------------------------------------
\begin{center}
{\LARGE \textbf{Saikou Ira — Supernatural Investigator}}\\[4pt]
{\large Kahfagian Marine Captain turned Occult Detective of The Witness}
\end{center}

% -------------------------------------------------------------
% PROFILE
% -------------------------------------------------------------
\section*{Profile}
\begin{tabular}{@{}ll}
\textbf{Name:} & Saikou Ira \\
\textbf{Concept:} & Kahfagian Marine Captain turned Supernatural Investigator \\
\textbf{Tier:} & III (68 XP) \\
\textbf{Affiliation:} & Kahfagian Navy, The Witness, The Empty Temple \\
\textbf{Familiar:} & Nyx — spirit cat, bound through mirror rite
\end{tabular}

\smallskip
\noindent
\emph{Main Patron: The Witness, Keeper of the Unblinking Truth.}  
Its symbol is a silver mirror set in obsidian: always clear of dust, never reflecting the caster.  
Those who serve The Witness walk a razor’s edge between revelation and ruin.

% -------------------------------------------------------------
% ATTRIBUTES
% -------------------------------------------------------------
\section*{Attributes}
\begin{tabular}{@{}ll}
\textbf{Body}  & 2 — disciplined but scarred \\
\textbf{Wits}  & 4 — tactical intellect, forensic insight \\
\textbf{Spirit} & 3 — steeled will, supernatural focus \\
\textbf{Presence} & 3 — commanding voice, analytical calm
\end{tabular}

% -------------------------------------------------------------
% SKILLS
% -------------------------------------------------------------
\section*{Skills}
Investigation 4; Arcana 3; Command 2; Sway 3; Lore 2; Melee 2

% -------------------------------------------------------------
% TALENTS
% -------------------------------------------------------------
\section*{Talents}
\begin{itemize}
  \item \textbf{Caster’s Gift} — Freeform ritual casting guided by The Witness.
  \item \textbf{Thiasos (Nyx)} — Spirit Familiar capable of passing through reflections, sensing lies, and retrieving echoes of truth.
  \item \textbf{Codex} — The Witness’ Runekeeper compendium; each rune is a truth etched in reflection.
  \item \textbf{Patron’s Symbols} — Divination (Truth), Abjuration (Wards), Exorcism (Banishment), Kahfagian Heritage (Raéyn’s favor).
  \item \textbf{Lorekeeper} — Recall obscure knowledge without rolling; The Witness whispers forgotten details.
  \item \textbf{Exceptional Coordination} — One follower provides +4 Assist dice.
  \item \textbf{Iron Stomach} — Immune to mundane toxins; truth cannot be poisoned.
\end{itemize}

% -------------------------------------------------------------
% PATRON
% -------------------------------------------------------------
\section*{Patron — The Witness}
\textbf{Symbol:} A silver mirror on obsidian base.\\
\textbf{Domain:} Investigation, revelation, clarity through confrontation.\\
\textbf{Epithets:} The Unblinking Eye, The Mirror Without Reflection, Judge of Lies.\\
\textbf{Doctrine:} ``There is no mercy in truth, but neither is there falsehood in justice.''\\[6pt]

\textbf{Rites:}
\begin{itemize}
  \item \textbf{Rite of the Lingering Glimpse (Low)} — Preserve a fading truth long enough to analyze; requires a reflective surface.
  \item \textbf{Rite of Piercing Scrutiny (Low)} — Read falsehood from the eyes or aura of a speaker.
  \item \textbf{Rite of the Echoing Truth (Standard)} — Manifest spoken truths as visible runes that linger briefly.
  \item \textbf{Rite of the Immutable Record (Standard)} — Inscribe an event into mirror or lens; cannot be erased.
\end{itemize}

% -------------------------------------------------------------
% FAMILIAR
% -------------------------------------------------------------
\section*{Familiar — Nyx}
A sleek black cat with heterochromia.  
Bound spirit of The Witness, serving as mirror and messenger.

\textbf{Abilities:}
\begin{itemize}
  \item Detect auras and falsehoods.
  \item Slip through reflections and polished metal.
  \item Telepathic link to Saikou; transfers whispers and visions.
  \item Retrieve ``echoes of truth'' from a crime scene, granting Saikou +1 die on the next Investigation roll.
\end{itemize}

\emph{Personality:} Sarcastic, watchful, unerring loyalty to the mirror law.  
\emph{Bond:} ``Truth hides in the dark, Captain. Let me bring the light that burns it clean.''

% -------------------------------------------------------------
% FOLLOWER — AMARI
% -------------------------------------------------------------
\section*{Follower — Amari, Kahfagian Marine Bodyguard}
\textbf{Tier:} III Follower (3–5 Cap)\\
\textbf{Role:} Protective shadow and shock infantry guard.

\textbf{Traits:}
Disciplined, vigilant, unshakably loyal to Saikou and the old Navy codes.

\textbf{Loadout:}
Reinforced boarding shields, barbed spear, storm-blue cloak, iron lantern.

\textbf{Special: Combat Boarding Specialist}
\begin{itemize}
  \item \textbf{Spear-and-Shield Blitz:} When Amari Charges a target within Near range, roll Melee. On a success, choose two:
  \begin{itemize}
    \item Knock the target Prone.
    \item Pin the target against cover or a wall.
    \item Disarm a held weapon.
    \item Advance Saikou into a better Position (e.g., Risky → Controlled).
  \end{itemize}
  On a Partial, choose one and Amari risks Level 1 Harm.
\end{itemize}

\textbf{Harm \& Morale:}
\begin{itemize}
  \item \textbf{Wounded:} -1 die to Athletics and Melee.
  \item \textbf{Broken Shield:} First time Level 2 Harm is taken, her shield splinters. She keeps fighting with spear alone.
  \item \textbf{Level 3 Harm:} Amari is incapacitated and Saikou gains +2 Boons to spend this scene.
  \item \textbf{If not recovered in 3 days:} Her Morale turns—she becomes obsessed with protecting Saikou, even at the cost of others.
\end{itemize}

\textbf{Bond Move — ``Not Again.''}\\
When Amari takes Harm \emph{for} Saikou, Saikou gains +1 die to the next Resolve or Investigation roll against fear, illusions, or mind-affecting magic.  
If Saikou Pushes that roll, add +2 dice instead.

\textbf{Ouroboros Tattoo (Debt Mark):}\\
A serpent eating its tail, inked in silver along Amari’s shoulder.  
Once per session, when Saikou gives an order that protects the innocent or reveals a hidden truth, Amari may refresh 1 Harm box or stand back up at Level 2 Harm for one more exchange before collapsing.

% -------------------------------------------------------------
% SIDEBAR: INVESTIGATION METHODS
% -------------------------------------------------------------
\section*{Sidebar: Investigation Techniques}
\begin{itemize}
  \item \textbf{The Canré Method (3 Moves Ahead):} Frame three hypotheses; sacrifice one each scene to force revelation. Spend 1 Boon to reposition socially after a risky reveal.
  \item \textbf{Lantern Pier Routine:} Log smell, sound, light, line before touching evidence. Once per case, recall one and gain +1 die.
  \item \textbf{Evidence Ladder:} Always convert one clue into an \emph{asset}—map-pin, oath, witness favor.
  \item \textbf{Ritual Forensics:} Pair Witness + Sealed Gate: reveal, then sanctify. Contamination attempts are Desperate.
  \item \textbf{Interview Triangulation:} Factual, motive, memory—each in separate locales. Contradictions = actionable leads.
  \item \textbf{Harbor Net (Kahfagia):} Spend 1 Boon to ``buy the tide'' for quiet access to a restricted quay.
  \item \textbf{Shadow Walk:} Move one block ahead; predict choices. Place a Waymark +1 die to ambush.
\end{itemize}

% -------------------------------------------------------------
% SIDEBAR: RECOVERY SCENE
% -------------------------------------------------------------
\section*{Recovery Scene — The Lantern and the Sea}
When tending Amari’s wounds, narrate a calm reflection.  
If Saikou speaks a hard truth aloud, remove 1 Level of Harm from Amari.  
If Saikou remains silent, Saikou instead clears 1 Stress.

% -------------------------------------------------------------
% SIDEBAR: PATRON REFERENCE
% -------------------------------------------------------------
\section*{Sidebar: Patron Interaction Reference}

\subsection*{The Witness — Truth/Revelation}
\textbf{Invoke:} First pass on a scene; deciding what’s real.\\
\textbf{Spend:} 1 Boon to stabilize a fragile vision.\\
\textbf{Conflict Sign:} Mirrors film over.\\
\textbf{Whisper:} ``Ask the question that costs a friend.''\\
\textbf{Overuse:} Cold judgment—-1 die to warmth/comfort until rest.

\subsection*{The Sealed Gate — Boundaries/Closure}
\textbf{Invoke:} When harm spreads: contagion, panic, spirits.\\
\textbf{Spend:} 1 Boon to anchor a ward to a \emph{name}.\\
\textbf{Conflict Sign:} Doors refuse to open twice.\\
\textbf{Overuse:} Collateral isolation.

\subsection*{Raéyn — Storms/Tides}
\textbf{Invoke:} Momentum shifts, maritime leads, crowd control.\\
\textbf{Spend:} 1 Boon to reposition the team without provoking.\\
\textbf{Conflict Sign:} Salt taste; floor tilts.\\
\textbf{Overuse:} Tempers flare—parley starts at worse Position.

\subsection*{The Traveler — Ways/Roads}
\textbf{Invoke:} Escapes, tails, pattern pursuit.\\
\textbf{Spend:} 1 Boon to turn a Waymark into a safe escape route.\\
\textbf{Conflict Sign:} Maps smudge at the margins.\\
\textbf{Overuse:} Timing slips—an ally arrives a scene late.

\medskip
\noindent
\emph{GM Note:} If two Patrons are invoked in the same exchange, consider a brief interference: a sign, a complication, or a minor cost. Success should still stand.

% -------------------------------------------------------------
% END
% -------------------------------------------------------------
\end{document}