% Fate's Edge SRD --- Combat (Section 08)

\section{Combat}

Combat uses the same core resolution loop as any other encounter: declare intent, set Position and DV, roll, apply outcome. This section provides specific rules for violence, movement, and tactical positioning.

\subsection{Rhythm of Battle}

Fate's Edge does not use fixed initiative. The GM spotlights characters based on the fiction, circumstances, and tension.

\begin{itemize}
  \item \textbf{Ambushers act first.} Targets respond after the opening exchange.
  \item \textbf{Spotlight shifts} mid-scene if fictionally appropriate (e.g., reacting to a falling ceiling, seizing a moment).
  \item \textbf{Players may suggest acting} when it makes sense in the fiction.
\end{itemize}

\subsection{Turn Structure}

Each character's turn consists of \textbf{1 Action} and \textbf{1 Move}. Actions and Moves may be taken in any order. Repeating the same Action is not allowed unless noted.

\subsubsection{Move}
Move up to one range band (Close $\rightarrow$ Near $\rightarrow$ Far, or vice versa). You may also use your Move to reposition within the same range band (e.g., from behind one tree to another at the same distance). This does not change your range band but may affect cover or line of sight.

\begin{itemize}
  \item \textbf{Stand:} Use your Move to stand from prone.
  \item \textbf{Dash (Action):} Gain an extra Move. You may move two bands total in your turn. Your next defense roll is Desperate.
\end{itemize}

\subsubsection{Disengage (Action)}
Prepare to retreat safely. After taking Disengage, your Move this turn does not provoke opportunity attacks. You cannot also Attack this turn.

\subsection{Range Bands}

\begin{tabular}{@{}llp{7cm}@{}}
\toprule
\textbf{Band} & \textbf{Default} & \textbf{Notes} \\
\midrule
Close & Touching & Melee reach; grapples, knives, short weapons. \\
Near & Same room/street segment & Thrown weapons, short bows, most magic. \\
Far & Distant (same site) & Long bows, line spells; needs time or route to close. \\
Absent & Off-screen & Outside scene frame; needs travel or cut to interact. \\
\bottomrule
\end{tabular}

\subsection{Melee Flag}

Two parties are ``in Melee'' when they are within Near and directly engaged.

\begin{itemize}
  \item \textbf{Enter Melee:} 1 Move from Near (Controlled if under fire).
  \item \textbf{Leave Melee:} 1 Move to break off. If threatened, you must take the \textbf{Disengage} action. Without Disengage, leaving Close range provokes a free opportunity attack.
\end{itemize}

\subsection{Opportunity Attacks}

Moving out of an enemy's Close range without using the Disengage action provokes one free melee attack from that enemy. The attack is resolved normally (Position usually Controlled) and does not consume the enemy's action for the round.

Additionally, the following actions provoke an opportunity attack if performed while engaged in Close with an enemy:
\begin{itemize}
  \item Making a ranged attack (bow, crossbow, thrown weapon).
  \item Casting a spell that does not have the \texttt{[TOUCH]} or \texttt{[WARD]} tag.
\end{itemize}

\subsection{Full Turn Actions}

\subsubsection{Charge}
Move up to two range bands toward an enemy (e.g., Far $\rightarrow$ Near $\rightarrow$ Close) and make one melee attack.
\begin{itemize}
  \item Costs 1 Fatigue.
  \item Your next defense is Desperate.
  \item Consumes both your Action and Move.
  \item Requires open ground (no difficult terrain) unless you have a talent that overrides this.
  \item If your target is hidden behind full cover, you cannot complete the Charge. You may move up to the cover and end your turn there; the attack is lost.
\end{itemize}

\subsubsection{Skirmish}
Move up to two range bands away from an enemy and make one ranged attack.
\begin{itemize}
  \item Costs 1 Fatigue.
  \item Your next defense is Desperate.
  \item Consumes both your Action and Move.
  \item Requires open ground.
  \item If you start your turn hidden, you may Skirmish: leave cover, move, shoot, and may end your turn in new cover if available. You are exposed during movement and the shot; opportunity attacks apply normally.
\end{itemize}

\subsection{Cover and Hiding}

\subsubsection{Cover}
\begin{itemize}
  \item \textbf{Partial cover:} +1 DV to target you.
  \item \textbf{Full cover:} You cannot be targeted directly.
\end{itemize}

\subsubsection{Hiding}
To hide, a character must be behind full cover or in heavy concealment (dense fog, darkness). Take the \textbf{Interact} action to hide, rolling Stealth (or appropriate skill) against the Perception of any observers.

\begin{itemize}
  \item Once hidden, you are not a valid target unless an attacker first uses an action to \textbf{Search} (Perception vs. your Stealth) or you reveal yourself.
  \item Revealing yourself by attacking or moving into the open ends the hidden state. You are exposed for the duration of the attack and may provoke opportunity attacks if you leave an enemy's Close range.
\end{itemize}

\subsection{Height Advantage}

Attacking from higher ground (e.g., a ledge, rooftop, or horse) grants +1 die or +1 Effect (GM's choice).

\subsection{Difficult Terrain}

\begin{itemize}
  \item Costs 2 Move actions to traverse one band (or a full action to Dash).
  \item Charge and Skirmish require open ground unless the character has the \textbf{Relentless} talent.
\end{itemize}

\subsection{Mounted Combat (Simplified)}

If you have a War Mount and the \textbf{Cavalier} talent (or equivalent):
\begin{itemize}
  \item \textbf{Charging} into melee from Far $\rightarrow$ Near: +2 dice.
  \item \textbf{Firing while disengaging} Near $\rightarrow$ Far: +2 dice.
\end{itemize}

The bonus applies only when movement is part of the action. Misses and Partials often cost mount Fatigue, equipment damage, or narrative complications.

\subsection{Combat Action Options}

In addition to standard actions (Attack, Disengage, Dash), characters may take:

\begin{itemize}
  \item \textbf{Shove (Action):} Opposed Body + Athletics vs. target's Body + Athletics. On success, push the target back one range band (Close $\rightarrow$ Near). The target suffers –1 Position on their next action.
  \item \textbf{Grapple (Action):} Opposed Body + Melee vs. target's Body + Athletics. On success, both characters are engaged (cannot move apart). The grappled target suffers –1 die on all actions. Breaking free requires a new opposed roll as an action.
  \item \textbf{Protect (Action):} Adopt a guarding stance. Choose a nearby ally; until your next turn you may intercept one hit on them and roll to resist it. On success, reduce or negate Harm; you take any fallout the GM assigns.
  \item \textbf{Setup (Action):} Create advantage for an ally. On a success (appropriate roll, DV set by GM), grant their next action +1 Position or step up Effect.
  \item \textbf{Defend (Action or Move):} Enter a defensive stance. Until your next turn, improve your Position by one step when resisting any attack or effect (or gain +1 die if already Dominant).
\end{itemize}

\subsection{Default Combat Assumptions}

To keep play moving, the GM may assume:
\begin{itemize}
  \item Default range = Near.
  \item Default Position = Controlled.
  \item Default Effect = Standard.
  \item Move one band per Move action unless terrain or fiction says otherwise.
\end{itemize}

\subsection{Example: Melee Exchange}

\textit{Thane charges a goblin archer.}

\begin{itemize}
  \item Thane declares \textbf{Charge} (Far $\rightarrow$ Near $\rightarrow$ Close). Costs 1 Fatigue.
  \item GM: ``Controlled, DV 3. On a Miss, the goblin gets a parting shot.''
  \item Thane rolls Body 3 + Melee 3 = 6 dice: 9, 7, 6, 4, 2, 1 $\rightarrow$ 3 successes vs DV 3. Clean Success.
  \item Thane deals his weapon's Harm (Medium weapon = Harm 2). The goblin is injured.
  \item Thane's next defense is Desperate due to Charge.
\end{itemize}

\subsection{Integration with Fatigue and Harm}

Combat frequently marks Fatigue (from Charge, Skirmish, and enemy abilities) and Harm (from successful attacks). See Section~\ref{sec:health-fatigue-harm} for recovery and the armor conversion table.

