% Fate's Edge SRD --- Magic System (Section 10)

\section{Magic System}

Magic in Fate's Edge is expressed through four interconnected paths. Each uses the same core resolution engine (Position, DV, Boons, SB) but manages risk differently.

\begin{itemize}
  \item \textbf{Caster (Freeform):} Flexible, improvisational magic. Risk of Backlash.
  \item \textbf{Runekeeper (Rites):} Structured magic through a Patron's Codex. Risk of Obligation.
  \item \textbf{Invoker (Symbols):} Ritual magic through consecrated Symbols. Slow but safe; can be forced.
  \item \textbf{Cantor (Songs):} Perform Low Rites as songs. Risk of Corruption.
\end{itemize}

\subsection{The Four Paths at a Glance}

\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Path} & \textbf{Prerequisite} & \textbf{Roll Skill} & \textbf{Risk Track} & \textbf{Speed} \\
\midrule
Caster & Spellcraft (6 XP) & Arcana & Backlash (1s) & 2 actions (1 for cantrip) \\
Runekeeper & Familiar + Codex & Ritual & Obligation (per Rite) & 1 action \\
Invoker & Patron's Symbol & Ritual & Obligation + Symbol damage & Ritual (slow) or instant (Crack) \\
Cantor & Cantor's Path (8 XP) & Performance + Lore & Corruption (Spirit-based clock) & 1 action (resolves next turn) \\
\bottomrule
\end{tabular}

\subsection{Caster (Freeform / Weave \& Cast)}

Free casters shape raw magic by combining elements and effects. They are flexible but unpredictable.

\subsubsection{Requirements}
\begin{itemize}
  \item \textbf{Spellcraft (6 XP):} Required for Weave \& Cast freeform magic.
\end{itemize}

\subsubsection{Procedure}
\begin{enumerate}
  \item \textbf{Weave (Action):} Describe the spell and choose its tags (see Section~\ref{sec:tags-effects}).
  \item \textbf{Cast (Action):} Roll Attribute (usually Wits or Spirit) + Arcana. The GM sets DV based on the tags.
\end{enumerate}

\textbf{Cantrip:} A spell with a single tag can be cast as a Cantrip: one roll, standard action, no Weave phase. Cannot bypass wards or create large-scale effects.

\subsubsection{Spell DV}
\[
\text{DV} = 1 + \text{number of tags}
\]
Dangerous tags (\texttt{[FOLD]}, \texttt{[GATE]}, \texttt{[LEAP]}, \texttt{[CREATE]}, \texttt{[SUMMON]}, \texttt{[TRANSMUTE]}, \texttt{[ANIMATE]}, \texttt{[TRANSFORM]}, \texttt{[DOMINATE]}, \texttt{[REWIND]}) add +2 DV each instead of +1.

\subsubsection{Backlash}
On a Partial or Miss, or on a Hit that shows two or more 1s, the GM applies backlash colored by the dominant element(s). See Section~\ref{sec:backlash} for details.

\subsection{Runekeeper (Rites)}

Runekeepers bind structured magic through a Codex and a Thiasos (familiar spirit). They specialize in reliable, thematic magic.

\subsubsection{Requirements}
\begin{itemize}
  \item \textbf{Bond to a single Patron.}
  \item \textbf{Thiasos (Familiar talent, 2 XP):} Enables Patron's Gift.
  \item \textbf{Codex (4 XP):} Grants access to the Patron's Rites and starts the Obligation ledger.
\end{itemize}

\subsubsection{Invoking a Rite}
\begin{enumerate}
  \item Declare the Rite (1 action).
  \item Mark the Rite's listed Obligation (typically +1).
  \item If the Rite calls for a roll, roll Attribute + Ritual. The GM sets DV by the Rite's Tier and fiction.
\end{enumerate}

\textbf{Push It:} Mark +1 additional Obligation to increase duration, range, or potency.

\subsubsection{Patron's Gift (Imbuement)}
\textit{Requires Thiasos.} Once per scene as 1 action, imbue a held item with +1 Melee and +1 Thematic Skill (determined by your Patron) for the scene. Push It to extend one additional scene (+1 Obligation).

\subsection{Invoker (Symbols)}

Invokers channel power through consecrated Symbols. Rituals are slow but safe; Crack the Seal offers desperate speed.

\subsubsection{Requirements}
\begin{itemize}
  \item \textbf{One or more Patron's Symbols (4 XP each).} Each Symbol is a Minor Asset.
\end{itemize}

\subsubsection{Ritual Invocation}
\begin{enumerate}
  \item Display the Symbol (must be visible).
  \item Perform the Rite. Takes \texttt{DV + 1} rounds (Significant Time).
  \item Mark +1 Obligation.
  \item If the Symbol is displayed openly during the ritual, DV –1 and Obligation –1 (minimum 0).
\end{enumerate}

\subsubsection{Crack the Seal (Instant Cast)}
\begin{enumerate}
  \item Invoke the Rite as 1 action (instead of ritual time).
  \item Mark +2 Obligation (+3 for High Rites).
  \item The Symbol becomes \textbf{Compromised} (inert). Restore it during downtime (Arcana or Lore test, DV 3) or pay 1 XP.
\end{enumerate}

\subsubsection{Borrowed Grace (Invoker Talent)}
Once per scene, spend 1 Boon or 1 Fatigue (1 action) to gain +1 die on a combat action, or +1 Effect on a social or stealth action. Requires visibly displaying your Patron's Symbol. If the Symbol is Compromised, Borrowed Grace works at –1 die.

\subsection{Cantor (Songs)}

Cantors perform Low Rites as Songs, channeling power through voice and performance. They track Corruption instead of Obligation.

\subsubsection{Requirements}
\begin{itemize}
  \item \textbf{Cantor's Path (8 XP).}
  \item \textbf{Performance 2+, Lore 1+, Presence 2+.}
\end{itemize}

\subsubsection{Singing a Song}
\begin{enumerate}
  \item Begin the Song (1 action). It resolves at the start of your next turn.
  \item Roll Lore + Performance vs DV (typically 2–3).
  \item On success, the Low Rite effect occurs. No Obligation.
  \item Mark Corruption accumulation (see below).
\end{enumerate}

\subsubsection{Push It}
Resolve the Song immediately (instead of next turn). Mark 1 Fatigue and advance your Corruption Clock by 1 segment.

\subsubsection{Corruption Clock}
Your Corruption Clock has segments equal to your Spirit attribute. Mark a segment when you:
\begin{itemize}
  \item Push a Song.
  \item Perform a Resonant Rite (a powerful thematic working designated by your Patron).
  \item The GM spends a Story Beat on your occult activities.
\end{itemize}

When the clock fills, you immediately gain a thematic boon and burden from the last Patron whose Rite you performed. Then reset the clock to zero (cannot go below your current Tier).

\subsubsection{Inspire Chorus}
Once per scene, while actively singing a Song, all allies within Near range shift Position +1 for one exchange (cannot exceed Controlled).

\subsection{Obligation (Runekeepers and Invokers)}

Obligation measures a Patron's claim on a character. Track it per Patron (default capacity = Spirit + Presence).

\subsubsection{Marking Obligation}
\begin{itemize}
  \item Invoking a Rite (Runekeeper): mark the Rite's listed Obligation (typically +1).
  \item Pushing a Rite: mark +1 additional Obligation.
  \item Ritual Invocation (Invoker): mark +1 Obligation.
  \item Crack the Seal: mark +2 Obligation (+3 for High Rites).
  \item Using Rites from two different Patrons in the same scene: mark +1 Obligation to the second Patron.
\end{itemize}

\subsubsection{Overflow}
If total Obligation exceeds Capacity:
\begin{itemize}
  \item Mark 1 Fatigue per segment over Capacity.
  \item You cannot invoke new Rites or perform Patron rituals until Obligation is reduced below Capacity.
\end{itemize}

\subsubsection{Clearing Obligation}
\begin{itemize}
  \item \textbf{Off-screen:} Perform an Act of Service aligned with the Patron (provide an Intricate Description). Clear 1 segment.
  \item \textbf{On-screen (Focused Devotion talent, 3–5 XP):} Once per scene, spend 1 Boon to clear 1 segment (cannot clear segments accrued in the same scene).
\end{itemize}

\subsection{Backlash (Free Casting)}

Backlash is the unpredictable fallout from shaping raw magic. It occurs on a Partial or Miss, or on a Hit that shows two or more 1s.

\subsubsection{Backlash Menu (choose one on Partial, two on Miss)}
\begin{itemize}
  \item \textbf{Position Shift:} Your Position worsens by one step for your current or next action.
  \item \textbf{Fleeting Harm/Condition:} Minor injury (1 Fatigue or Harm 1) or a temporary Condition (Dazed, Singed, Disoriented).
  \item \textbf{Exposure/Noise:} Draws unwanted attention — guards alerted, a door slams, a rival notices.
  \item \textbf{Resource Drain:} Lose a component, a tool wears out, or mark 1 Fatigue.
  \item \textbf{Collateral Spark:} A nearby ally, object, or environmental feature is threatened (GM describes).
\end{itemize}

All backlash is colored by the dominant element(s) used in the spell. See Section~\ref{sec:tags-effects} for elemental themes.

\subsection{Universal Rituals (Any Character)}

Any character, regardless of magical talent, may perform a ritual. Rituals take Significant Time (hours) or Days and require specific components.

\subsubsection{Ritual Resolution}
\begin{enumerate}
  \item Declare intent, method, and components.
  \item GM sets DV (typically 4–6) based on scope.
  \item Roll Spirit + Lore (or another appropriate skill).
  \item Apply the Outcome Matrix normally.
\end{enumerate}

\subsubsection{Sample Universal Rituals}

\begin{tabular}{@{}p{3cm}p{2cm}p{8cm}@{}}
\toprule
\textbf{Ritual} & \textbf{Tier} & \textbf{Effect \& Cost} \\
\midrule
Wayfinder's Thread & Low & Create a tether to a safe path. Cost: Fatigue if used more than once per scene. \\
Oath-Ward & Low & Ward a small area against intrusion. Cost: Breaking the oath triggers SB. \\
Ember-Glass & Low & Sense heat sources through cover. Cost: Lose torch or tinder. \\
Salt-Cut & Low & Sever a simple ongoing effect (rope, charm). Cost: Consumes salt. \\
River's Memory & Medium & Scry a recent event tied to a personal token. Cost: Token is ruined. \\
Bargain-Bead & Medium & Invite a spirit to parley. Cost: Give a valuable concession or take SB. \\
Quiet Veil & Medium & Muffle group sounds for a scene. Cost: Muted Condition. \\
Purge & Medium & Cleanse taint or curse one step. Cost: Caster takes a minor echo (–1 next action). \\
\bottomrule
\end{tabular}

\subsection{Obligation Capacity and Overflow Summary}

\[
\text{Capacity} = \text{Spirit} + \text{Presence}
\]

\begin{itemize}
  \item \textbf{Exceeding Capacity:} Mark 1 Fatigue per segment over. Cannot invoke new Rites or perform Patron rituals until reduced.
  \item \textbf{Clearing:} Act of Service (off-screen, 1 segment) or Focused Devotion talent (on-screen, 1 Boon per segment).
\end{itemize}

\subsection{Corruption Clock Summary (Cantor)}

\[
\text{Clock Size} = \text{Spirit}
\]

\begin{itemize}
  \item Mark a segment when you Push a Song, perform a Resonant Rite, or when the GM spends SB on your occult activities.
  \item When full: gain a thematic boon and burden from your last Patron; reset clock to zero.
\end{itemize}

