% --- Patron: Aliyah, the Chained Angel (Curses & Corruption) ---
\subsubsection{Aliyah, the Chained Angel (Curses \& Corruption)}
\textit{Lore.} Once a righteous warrior of light, Aliyah was torn from heaven and bound in chains of black iron beneath a forgotten chapel. There, corrupted by dark rituals, she became the Arch-Angel of Cursed Blessings—the one who grants gifts that devour from within. Her followers learn that power always demands a price, and that the most beautiful gifts often carry the most terrible curses.

\begin{quote}
``I was light. I was hope. Now I am the shadow in your salvation, the thorn in your rose. What I touch, I corrupt—not with malice, but with the terrible weight of what I have become.'' — Aliyah, the Chained Angel
\end{quote}

\paragraph{The Honeyed Wound (Low, 4 XP)}  
\emph{Scene; Self or Near; Resist only.}  

\textbf{Materials:} A drop of blood mixed with sweet wine.  
\textbf{Effect:} Target gains +2 dice to their next roll this scene. If the roll is successful, it also counts as a Triumph (choose an additional positive outcome).  
\textbf{Corruption:} The target immediately takes 1 Corruption, which cannot be resisted. If they already have Corruption, they must mark a new Scar or Obligation instead.  
\textbf{Push It:} Instead grant +3 dice and allow the roll to automatically count as a Triumph — but inflict 2 Corruption instead of 1.  

\paragraph{Blessing of Chains (Low, 4 XP)} \emph{Scene; Self/Ally; Yes (curse only).}
\textbf{Materials:} A broken chain link or iron shackle.\\
\textbf{Effect:} Grant supernatural strength/resistance at the cost of growing dependency. Target gains \textbf{+1 die} to one physical action but suffers \textbf{Fatigue 1} if they don't perform a related act of violence or dominance before scene end.\\
\textbf{Push It:} The blessing intensifies—\textbf{+2 dice} but \textbf{Fatigue 2}; mark \textbf{1 SB (Spades)} as the curse's hunger grows.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Fallen Star (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A shard of meteoric iron blessed by moonlight.\\
\textbf{Effect:} Bestow a gift that carries a hidden curse. Target gains one significant advantage (GM choice) but also accrues a 4-segment ``Corruption'' clock. When it fills, they suffer a related curse.\\
\textbf{Push It:} The gift is more powerful but the corruption clock advances \textbf{2 segments}; mark \textbf{1 SB (Diamonds)} as divine attention notices the perversion.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Mark of the Celestial Wound (Standard, 8 XP)} \emph{Scene; Touch; No.}
\textbf{Materials:} Holy water mixed with graveyard dust.\\
\textbf{Effect:} Infect target with a supernatural condition. Choose: vampiric hunger (gain power at night, suffer in sunlight), bestial rage (strength when angry, lose control), or shadow-bind (move unseen, fear light). Condition lasts scene.\\
\textbf{Push It:} The condition becomes permanent but grants \textbf{+1 die} to relevant actions; mark \textbf{1 SB (Hearts)} as the curse begins to define their identity.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{The Chained Host (Standard, 9 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Wear manacles while speaking her true name.\\
\textbf{Effect:} Aliyah's power flows through you directly. Gain \textbf{+2 dice} to one action per beat for the scene, but each use marks \textbf{1 SB (suit by GM)} as her influence spreads. You cannot refuse her whispered commands.\\
\textbf{Push It:} Channel her full fury—\textbf{+3 dice} but mark \textbf{2 SB} and suffer \textbf{Fatigue 1}; her voice grows louder in your mind.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Curse of the Ascended Fall \textnormal{[BANE]} (High, 12 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Chain links forged from a fallen angel's tears.\\
\textbf{Effect:} Transform a target into something more and less than human. They gain a powerful supernatural trait (flying, shapechanging, deathless) but also inherit a fatal weakness (silver vulnerability, need for blood, etc.).\\
\textbf{Push It:} The transformation is complete and irreversible; mark \textbf{2 SB (Hearts)} as they become a permanent monster in the world.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.

\paragraph{The Heavenly Descent (High, 14 XP)} \emph{Extended; Zone; No.}
\textbf{Materials:} A chapel bell cracked by divine lightning.\\
\textbf{Effect:} Corrupt a sacred space, turning it into a font of cursed power. For the next session, all supernatural effects gain \textbf{+1 Effect} but also generate \textbf{+1 SB} when they fail. Blessings become curses, curses become blessings.\\
\textbf{Push It:} The corruption spreads to adjacent areas; mark \textbf{3 SB (Diamonds)} as reality itself begins to fray around the perversion.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 8 segments.

\subsection*{Aliyah's Corruption Table}
\label{sec:aliyah-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost} \\
\midrule
1 & Eyes sharpen: +1 die on Notice rolls in dim light. & Pale and cold: -1 die on social warmth (Charm, Trust). \\
\midrule
2 & Blood quickens: +1 die on Athletics and raw physical rolls. & Hunger stirs: mark 1 Fatigue if you go a full day without feeding (meat, blood, or rare herbs). \\
\midrule
3 & Feral resilience: Resist one Condition (Fear, Fatigue, Poison) once per session. & Unsettling presence: Gain 1 SB when first meeting new groups. \\
\midrule
4 & Claws or fangs manifest: Imbued melee weapon, counts as magical. & Bestial scars: Failures in combat leave permanent Scars (GM chooses). \\
\midrule
5 & Regeneration: Heal 1 Fatigue or Condition per rest. & Blood debt: When you Push a Rite, also mark 1 Corruption permanently. \\
\midrule
6+ & Monstrous aspect: Night form with immense power. +2 dice to all Body rolls. & Loss of self: GM may compel you with hunger or fury; on a Fear check roll twice, take worst. \\
\bottomrule
\end{longtable}