% --- Patron: Aveh, the Rider Behind the Storm (Freedom & Erasure) ---
\subsection{Aveh, the Rider Behind the Storm (Freedom \& Erasure)}

\textit{Lore.} Aveh is the faceless rider at the horizon's edge, neither man nor woman, both and neither. The Ykrul call them \textit{Aveh}, a liminal spirit beyond hearth and vow, a storm-shadow that offers liberty at the cost of belonging. To travel with Aveh is to taste unchained freedom — and to risk being erased by the storm that follows. They are the patron of the outcast, the non-conforming, the genderless, and the forgotten. Aveh asks no oath and demands no loyalty, but every gift leaves a scar of absence.

\begin{quote}
``I am the track the storm devours. I am the freedom you fear and the silence you crave. To ride with me is to be unbound, and to be unbound is to vanish.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Freedom from Bonds:} escape, liberation, shedding constraints
  \item \textbf{Erasure:} becoming forgotten, removing traces, vanishing from memory
  \item \textbf{Liminal Identity:} non-conforming, genderless, between categories
  \item \textbf{Storm as Shelter:} the chaos that hides and the wind that carries away
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Aveh does not bind a willing servant — they befriend a creature as fleeting as the wind. Their \textbf{Thiasos} is a small beast or mythological being that embodies freedom: a storm-petrel that appears before gales, a gray fox that leaves no tracks, a shadow-hound pup whose fur shifts like cloud, or a raven that knows every crossroads. This creature requires no food, but it must be allowed to run free each day; if confined for more than a day, it becomes \textit{Compromised} (it refuses to return, or returns only to harry its keeper). The bond is one of mutual escape, not mastery.

The \textbf{Codex} of Aveh is never a book bound in one place. It may be a leather strap knotted with wind-signs, a set of chalked symbols that blow away after each use, or a slate etched with runes that fade in rain. Because Aveh's truths are written in absence, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week rewriting its contents from memory (or letting the wind rearrange them), or the Rites stored there become uncertain (treat as Compromised until restored).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Aveh may touch a held item (weapon, tool, or cloak) and whisper a word of unbinding. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Stealth} or \textit{Escape} rolls, and the first time each scene the bearer would provoke an opportunity attack, they may ignore it (they slip through the gap). After the scene, the user marks \textbf{+1 Obligation} as the wind's memory of them fades slightly.

\subsubsection*{Rites of Aveh}

\paragraph{Rite of the Unlatched Step (Basic, 5 XP)} \hfill \\
\emph{Scene; Self; Yes.} \\
\textbf{Tags:} \texttt{[LIBERATE]}, \texttt{[STEALTH]}, \texttt{[LIBERATE]} \\
\textbf{Materials:} A breath cast into the wind, or a line drawn in dust. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Storm-Step:} Slip free of notice or restraint. Gain \textbf{+1 die} on one escape or evasion roll this scene.
  \item \textbf{Exile's Banner:} Mark a target as ``outside.'' The target suffers \textbf{--1 die} on all social rolls for the scene.
\end{itemize}
\textbf{Push It:} (Storm-Step) Gain \textbf{+2 dice} instead, but mark \textbf{1 SB (Spades)} as your absence unsettles the scene. (Exile's Banner) The target is fully estranged; allies treat them as forgotten, but you immediately advance your Corruption timer by 1 segment. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Forgotten Road (Advanced, 9 XP)} \hfill \\
\emph{Scene; Zone; Yes.} \\
\textbf{Tags:} \texttt{[ERASURE]}, \texttt{[AREA]}, \texttt{[LIBERATE]} \\
\textbf{Materials:} A threshold or roadway under open sky, or stormwater rubbed into skin. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Erase the Road:} All attempts to follow you this session suffer \textbf{--2 dice}. In exchange, erase part of your own path: lose a detail (a memory, a name) chosen by the GM.
  \item \textbf{Rider's Mark:} A companion gains \textbf{+2 dice} to resist capture or coercion, but others forget them more easily.
\end{itemize}
\textbf{Push It:} (Erase the Road) Expand the effect to an entire group, but mark \textbf{1 SB (Clubs)} as reality warps around the erasure. (Rider's Mark) Grant \textbf{+3 dice} instead, but mark \textbf{1 SB (Diamonds)} as the Rider's storm distorts bonds. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Storm's Refuge (Master, 12 XP)} \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[INVISIBLE]}, \texttt{[AREA]}, \texttt{[ERASURE]}, \texttt{[LIBERATE]} \\
\textbf{Materials:} A banner or cloak raised in storm. \\
\textbf{Effect:} Your group is unseen or forgotten by pursuers for the scene. At the end, all present must mark \textbf{1 Corruption} or forget one bond (their choice). \\
\textbf{Push It:} Extend protection for an entire session, but mark \textbf{2 SB (Hearts)}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} 1 action; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 6 segments.

\paragraph{Rite of the Unremembered Name (Master, 13 XP)} \hfill \\
\emph{Extended; Self; Yes.} \\
\textbf{Tags:} \texttt{[ERASURE]}, \texttt{[TRANSFORM]}, \texttt{[FATE]}, \texttt{[LIBERATE]} \\
\textbf{Materials:} A circle of ash or dust drawn at a boundary, and a personal token (a lock of hair, a written name). \\
\textbf{Effect:} For one week, your name is erased from mundane records and casual memory. Those who knew you must succeed on a Resolve test (DV 4) to recall you. You gain \textbf{+2 dice} to all rolls to avoid detection, impersonate another, or pass unseen. At the end of the week, the erasure fades, but the GM may retain one lingering rumor or false memory. \\
\textbf{Push It:} The erasure becomes permanent — your original identity is gone. You cannot recover it except by divine intervention or a major quest. Mark \textbf{+3 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 8 segments.

\paragraph{Rite of the Horizon's Devour (Master, 15 XP)} \hfill \\
\emph{Extended; Zone; No.} \\
\textbf{Tags:} \texttt{[ERASURE]}, \texttt{[AREA]}, \texttt{[FATE]}, \texttt{[DEATH]} \\
\textbf{Materials:} A circle of ash or dust drawn at a boundary, and a storm that answers. \\
\textbf{Effect:} All thresholds in a zone collapse. Doors lead nowhere; borders dissolve; memory of the place becomes vague. In exchange, something of you is erased — your reflection, your name, or your presence in a specific person's memory (GM's choice). \\
\textbf{Push It:} The storm devours even time; mark \textbf{3 SB (Spades)} as history forgets you. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+3 Obligation}. \\
\emph{Obligation:} 8 segments.

\subsubsection*{Cantors \& Cults of Aveh}

\textbf{The Cantor's Song.} A Cantor of Aveh does not sing to be heard, but to be forgotten. Their art is the \textit{empty note}, the melody that leaves a silence where a memory once lived. They are sought by those who need to disappear — spies, exiles, the grief-stricken — and feared by those who want to be remembered.

\textbf{The Cult of the Unbannered.} Aveh has no temples, only crossroads. Their cults are small, mobile, and secretive. The Unbannered do not use names; they identify each other by a single knotted cord worn on the left wrist, the knot tied in a pattern that changes with every new moon. A ritual involves each member writing their name on a slip of paper, then passing the slips in a circle until no one remembers whose name they hold. They then burn the papers in a storm-lit brazier. The most devoted remove their own names from all records, wear featureless gray masks, and speak only through a designated ``voice'' who changes each season. Their creed: \textit{``We were never here.''} Outsiders who stumble into a meeting are offered a black feather and a choice: forget, or follow. Those who follow are never seen by their former lives again.

\subsubsection*{Witchcraft of Aveh (Lay / Threshold)}

A witch who walks with Aveh is a \textit{path-cutter}, a specialist in hidden roads and forgotten boundaries. Their power lies in omission: removing a door, erasing a trail, or unlatching the memory of a promise. Their tools are never hammers or chisels — they use whistles, feathers, and the quiet breath of a closed door.

\textbf{The Witch's Tool: The Whistle of the Unbound Road.} A small iron whistle, cold to the touch even in summer, shaped like a bird in flight. Once per session, when blown at a locked threshold (a door, a city gate, a sealed letter), the wind answers, and the lock opens as if it had never been closed. The whistle can also be used to ``call the storm'' — blown at a crossroads at midnight, it summons a thin, cold fog that obscures all tracks for an hour. The fog fades when the first cock crows.

\textbf{A Hedge Gift: Unlatched (Basic, 4 XP).} Once per scene, you may declare a mundane lock, latch, or simple binding to be ``unlatched'' and open freely. This works on doors, chains, shutters, and even a pursuer's grip, but not on magical wards.

\subsubsection*{Aveh's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost \\
1 & \textbf{Storm-sense:} +1 die to Notice rolls outdoors. & \textbf{Voice falters:} Others forget your words more easily. \\
2 & \textbf{Fleetness:} +1 die on escape or chase rolls. & \textbf{Shadow lags:} Mark 1 Fatigue when first seen in a scene. \\
3 & \textbf{Erasure's Mercy:} Once/session, negate 1 Harm or capture. & \textbf{Memory slips:} Allies forget one fact or bond tied to you. \\
4 & \textbf{Rider's Endurance:} Resist 1 Condition (Fear, Fatigue, Bind) once/session. & \textbf{Reflections falter:} You do not appear clearly in mirrors. \\
5 & \textbf{Storm-banner:} +1 die to inspire rebellion or defiance. & \textbf{Exile's weight:} Each Push of a Rite marks 1 permanent Corruption. \\
6+ & \textbf{Horizon's Claim:} You may ride unseen across any threshold. & \textbf{Self-erasure:} On social rolls, roll twice and take the worst result. \\
\end{tblr}

\subsection*{Playstyle Notes}
Aveh rewards those who embrace freedom, transience, and the courage to become unknown. Their followers are often outcasts, runaways, spies, or those who have rejected the bonds of family, nation, or gender. The cost is the slow loss of being remembered — friends forget your face, your name fades from records, and eventually, even your reflection may not recall you. Aveh does not punish; they merely offer, and the path of erasure is walked willingly. Ideal for players who enjoy themes of identity, liberation, and the bittersweet price of absolute freedom.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Freedom from Constraints:} escaping bonds of law, oath, and expectation
  \item \textbf{Erasure as Mercy:} the right to disappear, to be forgotten
  \item \textbf{Liminal Identity:} existing between categories, beyond gender and role
  \item \textbf{Storm as Shelter:} chaos that hides and wind that cleanses
  \item \textbf{Non-Conforming Patronage:} Aveh asks no conformity, only authenticity
\end{itemize}

% --- Monastic Tradition: Unbound Step (Aveh) ---
\begin{tcolorbox}[colback=black!5,colframe=gray!70!white,title=\textbf{Unbound Step (Aveh)}]
  \emph{I am the track the storm devours. To ride with me is to be unbound.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Wits 3+, Spirit 2+, Rite of the Unlatched Step or Cantor of Aveh.
  
  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} Chains cannot hold what the storm has erased. Unbound Step monks gain \textbf{+1 die} to resist any magical debt, geas, or contract that would restrict their movement or freedom of choice. If they succeed, the creditor's parchment briefly turns to ash — the clause was never written.
  
  \vspace{2mm}
  \noindent\textbf{Techniques (choose as Talents):}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.4,l]},
    rowhead = 1,
    row{odd} = {bg=gray!10},
    row{1} = {bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Storm-Step} (Basic, 6 XP) & Teleport to any unoccupied space within Near range, ignoring opportunity attacks. The space must be unobserved at the moment of arrival (the storm hides you). & 1 Fatigue & Wits 3 \\
  \textbf{Unlatched Step} (Basic, 5 XP) & Slip free of notice or restraint. Gain +1 die on one escape or evasion roll this scene. & 1 Boon & Spirit 2 \\
  \textbf{Forgotten Road} (Advanced, 9 XP) & All attempts to follow you this session suffer –2 dice. In exchange, you lose a detail (a memory, a name) chosen by the GM. & 1 Fatigue & Wits 2 \\
  \textbf{Rider's Mark} (Advanced, 8 XP) & A companion gains +2 dice to resist capture or coercion, but others forget them more easily. & 1 Boon & Spirit 3 \\
  \textbf{Storm's Refuge} (Master, 12 XP) & Your group is unseen or forgotten by pursuers for the scene. At the end, all present mark 1 Corruption or forget one bond (their choice). & 1 Fatigue & Tier II \\
  \textbf{Horizon's Devour} (Master, 15 XP) & Collapse a threshold: a door leads nowhere, a border dissolves. In exchange, something of you is erased — your reflection, your name, or your presence in a specific person's memory. & 1 Boon + 1 Fatigue & Tier III \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 13 XP):}
  \textit{Unremembered Name.} Once per session, erase your name from mundane records and casual memory for one scene. Gain +2 dice to avoid detection, impersonate another, or pass unseen. At the end of the scene, the erasure fades, but you may also teleport to any location you have ever visited (as long as you are willing to leave no trace of your passage). Gain a permanent \textbf{Debt Scar} (your name in any written record is slightly blurred) — but you may ignore one debt or obligation that was recorded under your old identity (the ledger no longer recognizes you).
\end{tcolorbox}