% --- Patron: The Confessor Beneath the Bell (Burden & Exchange) ---
\subsection{The Confessor Beneath the Bell (Burden \& Exchange)}

\textit{Lore.} The Confessor waits in places where guilt gathers – under church spires, in confessionals long abandoned, in the echo of unspoken sins. He does not judge, nor absolve. He \emph{takes}. Each confession is a transaction: your burden for his silence, your pain for his weight. Those who serve him walk with bowed heads, their mouths stitched by invisible threads, ringing faintly when they speak. To follow the Confessor is to understand that sin does not vanish – it migrates, carried from one soul to another until the bell tolls and the weight becomes unbearable.

The Confessor is the sin‑eater, the one who consumes guilt so that others may be light. But every meal leaves a stain. Those who take on the burdens of the guilty find that the guilt does not disappear – it waits, grows, and eventually demands to be passed on.

\begin{quote}
``Speak, and be freed. But understand – what you cast off, I will bear. And what I bear, you will one day inherit.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Sin Eating:} absorbing guilt, shame, and spiritual debt
  \item \textbf{Burden Transfer:} moving conditions, curses, and story beats from one creature to another
  \item \textbf{Confessional Exchange:} hearing secrets, trading silence for obligation
  \item \textbf{Weight of Ages:} the slow accumulation of sin that must eventually be released
\end{itemize}

\subsubsection*{Rites of the Confessor}

\paragraph{Rite of the Absorbed Sin (Low, 4 XP)} \hfill \\
\emph{Scene; Near/Touch; Yes.} \\
\textbf{Tags:} \texttt{[DEBT]}, \texttt{[TRANSFER]}, \texttt{[HEAL]}, \texttt{[FEAR]} \\
\textbf{Materials:} A hollow vessel or the ringing of a bell, or a red cord/shared drink. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Echo of Sin:} Absorb 1 Story Beat (Hearts or Diamonds) from an ally or NPC within Near range. Convert it into \textbf{+1 Obligation}. You know the emotional nature of the sin but not its source.
  \item \textbf{Burden Transference:} Transfer one Condition, Fatigue, or 1 SB from another target to yourself. You may attempt to reduce their Harm by 1 level; if successful, mark 1 Fatigue.
\end{itemize}
\textbf{Push It:} (Echo) Absorb 2 SB at once; mark \textbf{+1 Fatigue}. (Transfer) Split the burden among two allies; mark \textbf{1 SB (Hearts)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of Absolution by Ash (Low, 5 XP)} \hfill \\
\emph{Scene; Touch; Yes (sustain by prayer).} \\
\textbf{Tags:} \texttt{[PURIFY]}, \texttt{[DEBT]}, \texttt{[TRANSFER]}, \texttt{[FEAR]} \\
\textbf{Materials:} A pinch of consecrated ash or soot from a burned confession. \\
\textbf{Effect:} Lay hands upon a penitent and whisper a truth they cannot bear. Reduce their \textbf{Corruption by 1}. The sin transfers to you as \textit{Burdened Heart}: mark 1 SB (Hearts) or \textbf{+1 Obligation}. \\
\textbf{Push It:} Cleanse multiple targets within Close range (e.g., after a battle); each additional soul adds \textbf{+1 Obligation}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Sin-Eater's Pact (Standard, 8 XP)} \hfill \\
\emph{Scene; Touch; No.} \\
\textbf{Tags:} \texttt{[BOND]}, \texttt{[DEBT]}, \texttt{[TRANSFER]}, \texttt{[COMMAND]} \\
\textbf{Materials:} Shared confession whispered into an object or tattooed sigil. \\
\textbf{Effect:} Create a bond between you and the confessor; you may suffer 1 Fatigue to grant \textbf{+2 dice} to their Resolve test. If they fail, you suffer 1 Harm instead of them. \\
\textbf{Push It:} You may take both Harm and a Condition; mark \textbf{1 SB (Clubs)}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Echoed Confession (Standard, 7 XP)} \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[COMMAND]}, \texttt{[FEAR]}, \texttt{[TRUTH]}, \texttt{[DEBT]} \\
\textbf{Materials:} A small handbell or resonant surface. \\
\textbf{Effect:} Once/scene, replay the echo of a confession you have absorbed. Compel a target who shares its nature (Guilt, Regret, Betrayal) to act or reveal truth; DV = your Presence. \\
\textbf{Push It:} The echo manifests physically; mark \textbf{+1 Obligation} and \textbf{1 SB (Diamonds)}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Weight of Ages (High, 14 XP)} \hfill \\
\emph{Extended; Touch/Zone; No.} \\
\textbf{Tags:} \texttt{[DEBT]}, \texttt{[TRANSFER]}, \texttt{[FATE]}, \texttt{[AREA]}, \texttt{[FEAR]} \\
\textbf{Materials:} An ancient bell that has tolled for every execution in a kingdom, a chain forged from confession nails, the ashes of a martyr who died for silence. \\
\textbf{Effect:} Transfer accumulated guilt of generations to a single target. Target must make Spirit+Resolve (DV 5) or become overwhelmed by ancestral shame, suffering automatic Harm 2. Alternatively, release all burdens at once: clear all absorbed SB and Obligation. Every affected ally rolls Resolve (DV 4) or relives their confessed sin as hallucination. \\
\textbf{Push It:} (Transfer) The target's guilt infects their bloodline; mark \textbf{3 SB (Spades)}. (Release) Bind a spirit or soul within the toll; mark \textbf{1 SB (Spades)} and \textbf{+1 Obligation} permanently. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 8 segments.

% --- Sin-Eater Monastic (tcolorbox) ---
\begin{tcolorbox}[colback=black!5,colframe=gray!70!black,title=\textbf{Sin‑Eater Monastic (Order of the Unburdened Bell)}]
\emph{The bell tolls for thee – and I answer. Your sin becomes my silence. Your shame becomes my scar.}

\vspace{2mm}
\noindent\textbf{Prerequisites:} Spirit 3+, Resolve 2+, Rite of the Absorbed Sin or Rite of Absolution by Ash.

\vspace{2mm}
\noindent\textbf{Sin‑Eater Techniques (choose as Talents):}

\begin{tblr}{
  colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.3,l]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Technique & Effect & Cost & Req. \\
\textbf{Stitched Silence} & You may speak a lie without being detected by magical truth‑telling (e.g., Zone of Truth, Oath of Flame’s Sunlit Parley). The lie must be about a sin you have absorbed. & 1 Fatigue & Spirit 3 \\
\textbf{Bell’s Witness} & Once per scene, when you hear a secret confession, you may ring a small bell. The GM must answer one yes/no question about whether that confession was true. & 1 Boon & Tier II \\
\textbf{Burden Steady} & You may carry one additional Condition without penalty (e.g., you can be both Poisoned and Cursed without stacking penalties beyond the first). & Passive & Spirit 3, Body 2 \\
\textbf{Sin‑Eater’s Resilience} & When you take a Condition onto yourself from an ally, you may reduce the Fatigue cost by 1 (minimum 0). & Passive & Tier II \\
\textbf{Absolution’s Echo} & Once per session, after you absorb a Story Beat from an ally, you may immediately clear one minor Condition (Fear, Shaken, Guilty) from that ally. & Free & Tier III \\
\textbf{Confessor’s Toll} & When you use Rite of the Echoed Confession, the target suffers –1 die to resist if they have previously confessed to you (even indirectly). & Passive & Presence 3, Tier III \\
\end{tblr}

\vspace{2mm}
\noindent\textbf{Master Technique (Epic Talent, 12 XP):}
\textit{The Unburdened Bell.} Once per session, you may ring a great bell (real or imagined) and declare a single sin you have absorbed to be \textit{forgiven}. Remove that sin’s associated Story Beat or Obligation from yourself completely (it does not transfer). For the remainder of the session, you gain +2 dice to all rolls made to protect or absolve others. When the session ends, you gain 1 permanent Corruption – the Confessor remembers that you cheated the cycle. You cannot use this talent again until you have taken a major sin onto yourself (GM discretion).
\end{tcolorbox}

\subsubsection*{Cantors \& Cults of the Confessor}

\textbf{The Cantor's Song.} A Cantor of the Confessor sings in bells, not words – their voice strikes the air like a clapper on bronze. They perform \textit{intonations of release}, melodies that invite confession from unwilling throats and then swallow it whole. Their songs are used by torturers who wish to be merciful, and by priests who cannot bear the weight of what they hear.

\textbf{The Cult of the Silent Bell.} The Confessor's cults gather in places where guilt is already thick: courthouses after sentencing, battlefields the morning after, deathbeds where the dying strain to speak. They wear grey hoods and carry small handbells that they ring exactly once at the start of a confession, and once when the burden has been transferred. Their creed is simple: \textit{"We do not forgive. We carry. When we can no longer carry, we pass the load to the one beneath the bell."}

\subsubsection*{Witchcraft of the Confessor (Lay / Threshold)}

A witch who walks with the Confessor is a \textit{bell‑wright} or \textit{burden‑stitcher} – a crafter of thresholds that listen and containers that hold. Their work is often indistinguishable from craft: forging a bell that echoes only the unspoken, carving a confessional that remembers every whisper, stitching a cord that ties sin to stone. Their tools are not hammers; they use bells, cords, and the patient resonance of a struck bronze note.

\textbf{The Witch’s Tool: The Bell of Unspoken Burdens.} A small handbell of tarnished bronze, its clapper carved from a knucklebone. Once per session, the witch may ring the bell once. For the next scene, they can hear any whispered confession made within Near range, even if the speaker intended it to be silent. The bell also allows them to “absorb” a single SB from an ally as a free action (once per scene), but doing so marks 1 Fatigue. The bell’s tone is mournful and lingers longer than it should.

\textbf{A Hedge Gift: Bell of Borrowed Silence (2 XP).} Once per scene, you may tap a small handbell and absorb one mundane sound (a scream, a lie, a name spoken in anger) into the bell's resonance. For the remainder of the scene, you may release that sound as a distraction or a weapon (\emph{+1 die to Intimidation or Deception}).

\subsubsection*{Confessor's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Confessor's Ear:} +1 die to Insight when someone speaks a hidden fear or guilt. & \textbf{Ringing Silence:} You hear phantom bells; checks that require perfect silence suffer --1 die. \\
2 & \textbf{Burden Bearer:} Once/scene, convert 1 SB from an ally into 1 Fatigue for yourself. & \textbf{Stitched Mouth:} You cannot speak a lie about a sin you have absorbed; if you try, mark 1 Fatigue. \\
3 & \textbf{Weight of Shadows:} Once/session, cause a target who has confessed to you to hesitate (lose Initiative or act at Desperate Position). & \textbf{Shared Agony:} When you transfer a Condition to yourself, you also transfer its outward sign (a scar, a tremor). \\
4 & \textbf{The Bell Answers:} Once/session, ring a bell and name a sin; all who share that sin within Near must test Resolve (DV 3) or admit it aloud. & \textbf{Echoing Skin:} Your body remembers every burden carried; touching you inflicts 1 Fatigue on the other party. \\
5 & \textbf{Absolution by Proxy:} Once/session, fully clear a Minor Condition from an ally by taking it onto yourself. & \textbf{Permanent Toll:} Each time you use this benefit, mark +1 permanent Obligation to the Confessor. \\
6+ & \textbf{The Weight of Ages:} You may transfer a Major Condition (or Harm 2) from an ally to yourself without a roll. & \textbf{Bell's Heir:} The Confessor whispers that you will one day take his place – or break. Mark +3 Obligation. \\
\end{tblr}

\subsection*{Playstyle Notes}
The Confessor is the patron of sin‑eaters – those who take on the guilt of others so that the guilty may walk free. His followers gain power over burden transfer, confession, and the slow accumulation of spiritual debt. The cost is that every sin absorbed leaves a mark, and the weight of ages grows heavier with each secret carried. Ideal for players who enjoy redemption arcs, martyrdom, and the moral complexity of carrying others' sins.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Sin Eating:} consuming guilt, shame, and spiritual debt
  \item \textbf{Burden Transfer:} moving conditions, curses, and story beats
  \item \textbf{Confessional Exchange:} secrets as currency, silence as payment
  \item \textbf{Weight of Ages:} the slow accumulation of sin that must eventually be released
  \item \textbf{The Bell as Witness:} the sound that marks the beginning and end of every transaction
\end{itemize}