% −−− Patron: The Gallow's Bell (Justice & Judgment) −−−
\subsection{The Gallow's Bell (Justice & Judgment)}

\textit{Lore.} The Bell does not rage; it tolls. Cold and impartial, it measures all accounts in time. Its keepers are silent arbiters who weigh deeds against consequence, not out of anger but out of inevitability. To call upon the Bell is to bind oneself to the gravity of truth, where even silence is judged, and every oath leaves a resonance in iron.

Among the peoples of Fate's Edge, the Gallow's Bell is the patron of judges, notaries, and those who speak final words. In Ecktoria, the Bell is invoked by Censor's clerks who stamp death warrants. In Viterra, hedge‑courts ring a small iron bell to seal a verdict. In Linn, a broken bell at a Thing‑holm is a sign that a judgment was unjust and the spirits are still listening. Wherever a reckoning must be witnessed, the Gallow's Bell tolls.

\begin{quote}
What is broken must be mended, what is owed must be paid. The Bell remembers all reckonings.
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  ℹtem \textbf{Weighing of Deeds:} measuring action against consequence, debt against payment
  ℹtem \textbf{Solemn Judgment:} verdicts that cannot be appealed, sentences that must be carried out
  ℹtem \textbf{Compelled Truth:} forcing honesty under threat of the Bell's toll
  ℹtem \textbf{Iron Witness:} the Bell as an impartial recorder of oaths, lies, and reckonings
\end{itemize}

⫕section*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to the Gallow's Bell binds a creature of quiet judgment. Their \textbf{Thiasos} is a small beast or mythological being that embodies impartial witness: a grey mouse that lives in courthouses and never cheeps false testimony, a small brass‑bell frog whose croak sounds like a distant chime, a black‑and‑white cat that sits on the judge's bench and stares at those who lie, or a mechanical songbird that rings a tiny bell each time a promise is broken. This creature requires no food, but it must witness a formal promise or a judgment each day. If neglected for a week, it becomes \textit{Compromised} (it rings at the wrong moment or remains silent when truth is spoken, imposing –1 die on all social rolls involving oaths or honesty for its bearer).

The \textbf{Codex} of the Gallow's Bell is never a book of spells alone. It may be a court ledger bound in iron, each page sealed with a drop of wax for every case heard; a set of bronze bells, each inscribed with a precedent; or a slate tablet that records the last verdict as a faint, glowing chime. Because the Bell’s judgments are final, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week ringing each bell or re‑reading each precedent, or the Rites stored there become uncertain (treat as Compromised until re‑rung).

⫕section*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of the Gallow's Bell may touch a held item (weapon, shield, or tool) and speak a single word of judgment. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Command} or \textit{Insight} rolls when delivering a verdict, enforcing an oath, or uncovering a lie. Additionally, the first time each scene the bearer would be deceived by a lie, they may reroll the opposed check (the Bell’s resonance warns them). After the scene, the user marks \textbf{+1 Obligation} as the weight of the Bell’s attention lingers on their soul.

⫕section*{Rites of the Gallow's Bell}

\paragraph{Rite of the Measured Debt (Low, 4 XP)} ℎfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[COMMAND]} \\
\textbf{Materials:} A pair of scales balanced with tokens from both sides. \\
\textbf{Effect:} Establish a temporary accord. Both parties suffer −−1 die if they break it first. You gain +1 die to enforce compliance. \\
\textbf{Push It:} The accord is mystically weighted; breach inflicts \textbf{1 SB (Hearts)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Weighed Heart (Low, 5 XP)} ℎfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[SENSE]} \\
\textbf{Materials:} A small brass scale touched briefly to the chest. \\
\textbf{Effect:} Sense if the target acts against their nature or oath. Gain +1 die when pressing them. \\
\textbf{Push It:} Target must test Resolve (DV 3) or disclose a hidden conflict. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Balanced Scales (Standard, 8 XP)} ℎfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[DEBT]}, \texttt{[BOND]}, \texttt{[TRANSFER]} \\
\textbf{Materials:} A set of scales inscribed with runes of parity. \\
\textbf{Effect:} Exchange a burden between two willing parties (Harm for Fatigue, Debt for Favor, etc.). Both gain +1 die to cooperate. \\
\textbf{Push It:} May compel an unwilling exchange with contested Command + Wits. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Judge's Eye (Standard, 7 XP)} ℎfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[FEAR]} \\
\textbf{Materials:} A black hood worn in silence for one minute. \\
\textbf{Effect:} Detect lies within Near range; +2 dice to Insight. Liars suffer −−1 die to maintain their falsehood. \\
\textbf{Push It:} All deception is laid bare for the scene, but mark Exposure +1. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Final Reckoning (High, 13 XP)} ℎfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[COMMAND]}, \texttt{[TRUTH]}, \texttt{[FEAR]} \\
\textbf{Materials:} A circle of iron bells, each etched with nameless runes. \\
\textbf{Effect:} The Bell tolls through you. All present feel compelled to name a debt or wrongdoing. Those who lie suffer Harm 2; those who speak true gain +2 dice to persuasion for the scene. \\
\textbf{Push It:} The Reckoning manifests as spectral echoes of past wrongs −− liars automatically suffer narrative punishment (Keeper decides). \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Great Adjudication (High, 14 XP)} ℎfill \\
\emph{Extended; Zone; No.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[COMMAND]}, \texttt{[SUMMON]}, \texttt{[FATE]} \\
\textbf{Materials:} A consecrated gavel or a great bell struck three times. \\
\textbf{Effect:} Convene an unseen tribunal. Shadows of former judges and wronged souls gather to preside. For the next session, disputes within the zone are judged formally: +2 dice to Command when speaking as arbiter, and honest testimony gains +1 die. \\
\textbf{Push It:} The tribunal's verdict echoes beyond the zone, affecting one major conflict elsewhere. Mark \textbf{2 SB (Hearts)} as higher powers of judgment take notice. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 8 segments.

⫕section*{Cantors & Cults of the Gallow's Bell}

\textbf{The Cantor's Song.} A Cantor of the Bell sings not with melody, but with \textit{rhythm of measure} – a slow, tolling chant that mimics the swinging of a clapper. Their voice carries the weight of finality; one note can seal a verdict, another can suspend a sentence. They are employed by courts to “sing the sentence” into being, and feared by those who know that a Cantor's silence after a verdict is the most damning judgment of all.

\textbf{The Cult of the Unstruck Bell.} The Bell's cults are austere, gathered in bare stone chambers where a single great bell hangs unstruck. Before any ritual, the cultists argue the case – who owes what, who has paid, who must be reminded. Only when consensus is reached does someone strike the bell, and its tolling \textit{makes the judgment real}. Their creed is carved into the bell's lip: \textit{“We do not create debts. We reveal them. The Bell remembers what the living choose to forget.”}

⫕section*{Witchcraft of the Gallow's Bell (Monastic / Notarial)}

A witch who walks with the Bell is often a \textit{notary of the silent court}, a \textit{bell‑tolling clerk}, or a \textit{judge's scribe} – a nun, monk, or lay scholar who tends the records of obligation, copies the judgments of the Bell, and keeps the weights of the scales true. Their magic is the art of measurement and witness: sealing verdicts with leaden seals that glow when a debt is called, scribing judgment scrolls that burn with the truth of a confession, or maintaining the bell‑roster that notes every name that owes an accounting. In Aeler, a True‑Mason might carve a “weigh‑stone” into a courtroom floor, so that every testimony given upon it tips an unseen scale. In Viterra, a hedge‑notary might stitch a ledger whose entries tighten like knots when a payment is overdue. Their tools are not hammers but seals, quills, and the small brass bells that ring in the ears of the perjured.

\textbf{The Witch’s Tool: The Seal of Unspoken Weights.} A small iron seal, cool to the touch, stamped with a bell that has no clapper. Once per session, the witch may press the seal onto a document, a threshold, or a debtor’s palm. For the next day, any creature that knowingly breaks a promise recorded in that document (or sworn on that threshold) suffers a sharp sting (1 Fatigue) as the seal “remembers” the weight of their word. The seal can also be used to “strike” a bell: tapping it against any iron bell forces the bell to ring once, even if broken or without a clapper, and the ring can be heard only by those who owe a debt in the current scene.

\textbf{A Hedge Gift: Bell‑Toll of the Hidden Debt (2 XP).} Once per scene, the witch may whisper the name of a creature within Near range while tapping any ringing surface (a bell, a cup, a blade). For the next minute, that creature hears a faint, distant bell toll each time they attempt to lie or conceal an obligation. The GM gains \textbf{1 SB} (the Bell records the strain), and the creature suffers \textbf{−−1 die} on Deception rolls during that time. The witch may also ask the GM one yes/no question about a hidden debt owed by that creature; the GM must answer truthfully, though cryptically.

⫕section*{Gallow's Bell Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Judge's Intuition:} +1 die to Insight when weighing truth. & Must point out falsehoods when noticed, regardless of tact. \\
2 & \textbf{Quiet Authority:} Once/scene, treat a failed Command as success; mark 1 SB (Hearts). & Cannot remain neutral in disputes; indecision costs 1 Fatigue. \\
3 & \textbf{Scales of Balance:} Once/session, enforce an exchange of burdens. & Compelled toward fairness even when it hinders you. \\
4 & \textbf{Bell's Resonance:} +2 dice when calling for judgment or demanding restitution. & Suffer 1 Fatigue if wrongdoing is ignored. \\
5 & \textbf{Reckoner's Call:} Once/session, declare a “reckoning moment” – truth must surface or consequence falls. & Cannot ignore pleas for justice without marking 1 SB (Spades). \\
6+ & \textbf{Final Arbiter:} Once/session, render an absolute decree; all within the scene must obey or suffer consequence (Harm 1, +1 Obligation). & Mark +2 Obligation; the Bell demands you bear the weight of enforcement. Your voice tolls like a bell when you speak judgment – those who hear you cannot interrupt until your sentence is complete. \\
\end{tblr}

\subsection*{Playstyle Notes}
The Gallow's Bell rewards those who speak truth, enforce oaths, and deliver justice without cruelty. Its followers are not executioners – they are arbiters who ensure that every debt is weighed and every consequence is fair. The Bell does not punish the desperate; it punishes the deceiver. Intrusions often take the form of an unexpected witness (a bell that rings at the wrong moment, a forgotten promise resurfacing) or a demand to judge a dispute that cannot be settled by silence. Ideal for characters who serve as magistrates, justiciars, or honest mediators in a world of broken oaths.

\noindent\textbf{Emphasises}
\begin{itemize}
  ℹtem \textbf{Solemn Judgment:} verdicts that cannot be appealed
  ℹtem \textbf{Compelled Truth:} forcing honesty under the threat of the Bell
  ℹtem \textbf{Weighing of Deeds:} measuring action against consequence
  ℹtem \textbf{Iron Witness:} the Bell as impartial recorder
  ℹtem \textbf{The Toll of Silence:} what is not said is also judged
\end{itemize}