% --- Patron: Grimmir, the Old Man of the Forest (Primal Mystery & Seasonal Sacrifice) ---
\subsection{Grimmir, the Old Man of the Forest (Primal Mystery \& Seasonal Sacrifice)}

\textit{Lore.} Grimmir is the Old Man of the Forest, the presence that watches from the shadow of the oldest oak, the cold breath on the back of the neck when the path grows dark. He is not kind, nor is he cruel – he is \textit{wild}, and the wild does not negotiate. It demands, and it takes. His followers learn to read the language of blood on snow, the geometry of root and stone, and the weight of a heart offered at the turning of the year. Grimmir does not teach harmony; he teaches respect through fear, and fear through sacrifice.

Those who walk his path become keepers of the old ways: druids who know when a grove must be fed, rangers who leave the first kill for the soil, and hedge‑folk who whisper the names of spirits not to be spoken aloud. Their magic is not the gentle coaxing of garden plants but the raw, patient hunger of the forest floor – a hunger that swallows bones and grows mushrooms in the shape of the dead.

Among the peoples of Fate's Edge, Grimmir is called by many names. In Aeler, he is the \textit{Root‑Father}, a patron of deep fungus and stone‑lichen whose shrines are carved into living rock. In Ykrul, he is the \textit{Sky‑Wolf}, a spirit of the steppe who herds the seasons and grants omens. In Aelaerem, he is the \textit{Thorn‑Father}, keeper of hedgerow magic, honoured with red thread tied around roadside stones. Wherever the wild presses against the civilised, Grimmir waits.

\begin{quote}
“The forest does not ask. It takes. The deer does not weep for the wolf, and the wolf does not beg the earth. Give what the season demands, or the season will take it anyway.”
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Seasonal Sacrifice:} the old ways of blood and offering, the turning wheel that demands payment
  \item \textbf{Folk Horror:} the terror of the unseen, the ritual that must be performed, the boundary that must not be crossed
  \item \textbf{Primal Mystery:} knowledge hidden in root and stone, the language of omens and portents
  \item \textbf{Growth from Death:} the forest floor as a stomach; decay as the engine of new life
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Grimmir does not bind a familiar of hearth and home. Their \textbf{Thiasos} is a small beast or mythological creature of the wild wood: a pine marten with eyes like amber, a raven that knows the migration routes of every bird, a fox with a single white paw that leaves no prints in snow, a toad that hibernates in rotten logs and sings at the first thaw, or a stag‑calf with antlers budding like spring shoots. This creature requires no food in the usual sense, but it must be allowed to roam freely in unspoiled nature each day. If confined for more than a day, it becomes \textit{Compromised} (it scratches at doors, refuses to eat, and its coat grows dull; its bearer loses the benefit of \textit{Seasonal Sense} until it is released).

The \textbf{Codex} of Grimmir is never a book of paper and ink. It may be a slab of slate carved with seasonal runes that frost over in winter, a bundle of bark scrolls sealed with pine resin, a set of knotted cords woven from the hair of a wild boar, or the hollowed horn of a stag filled with charcoal and ash. Because the wild resists being fixed in writing, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week in a natural place, re‑carving or re‑knotting the Rites by memory, or the symbols fade (treat as Compromised until restored under open sky).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Grimmir may touch a held item (weapon, tool, or garment) and whisper a word of the season. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Survival} or \textit{Stealth} rolls in natural terrain, and the first time each scene the bearer would be tracked by mundane means, the tracker suffers –1 die (the forest hides their trail). After the scene, the user marks \textbf{+1 Obligation} as the wild’s claim on them deepens.

\subsubsection*{Rites of Grimmir}

\paragraph{Rite of the Blood Offering (Low, 4 XP)} \hfill \\
\emph{Scene; Touch; Yes.} \\
\textbf{Tags:} \texttt{[SACRIFICE]}, \texttt{[NATURE]}, \texttt{[STEALTH]} \\
\textbf{Materials:} A drop of your own blood (1 Fatigue) or the blood of a freshly killed animal. \\
\textbf{Effect:} The forest accepts your offering. For the remainder of the scene, you gain \textbf{+1 die} to all \textit{Survival} and \textit{Stealth} rolls in natural terrain. Additionally, you may ask the GM one yes/no question about a threat hidden in the wilderness (e.g., “Is there a predator watching us?”). \\
\textbf{Push It:} The offering is accepted with hunger – you gain \textbf{+2 dice} instead, but you must leave a visible token (a hair, a scrap of clothing) on the ground. That token can be used by the GM later as a sympathetic link (spend 1 SB to have an animal or spirit track you). \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Speaking Roots (Low, 5 XP)} \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[NATURE]},  \\
\textbf{Materials:} Soil from a grave or a crossroads, held in the mouth. \\
\textbf{Effect:} You listen to the memory of the land. Gain \textbf{+2 dice} to \textit{Investigation} or \textit{Lore} when seeking to understand what happened in a natural place – a death, a ritual, a violation of the wild. The GM must give you a truthful but cryptic answer. \\
\textbf{Push It:} The roots speak too clearly. You see a vision of the event, but it leaves you shaken: mark \textbf{Fatigue 1} and the GM gains \textbf{1 SB}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Heart’s Meal (Standard, 9 XP)} \hfill \\
\emph{Instant; Touch; No.} \\
\textbf{Tags:} \texttt{[CONSUME]}, \texttt{[HEAL]}, \texttt{[BOOST]}, \texttt{[FEAR]} \\
\textbf{Materials:} The fresh heart of an animal (or a sentient creature, if you dare), still warm. \\
\textbf{Effect:} You consume the heart raw. Gain one of the following:
\begin{itemize}
  \item Heal \textbf{1 Harm} (the heart’s vitality transfers to you).
  \item Gain a \textbf{+2 die} bonus on your next physical action (the heart’s strength).
  \item Learn a secret the creature knew in life (GM discretion; the heart whispers as you eat).
\end{itemize}
After the rite, you are \textit{marked} – animals within Near range are nervous around you (disadvantage on animal handling) for one scene. \\
\textbf{Push It:} Eat the heart of a thinking creature. The benefit is doubled, but you gain a \textit{Hunger} Condition (must eat another heart within a month or suffer 1 Fatigue per week). The GM gains \textbf{2 SB}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Twisted Path (Standard, 7 XP)} \hfill \\
\emph{Scene; Near; Yes.} \\
\textbf{Tags:} \texttt{[CURSE]}, \texttt{[TERRAIN]}, \texttt{[FEAR]} \\
\textbf{Materials:} A knotted cord woven from thorns and hair. \\
\textbf{Effect:} The forest bends against your enemies. Choose a single target within Near. For the remainder of the scene, all their navigation, tracking, and movement rolls suffer \textbf{–1 die} as roots trip them, branches block their way, and the path turns subtly wrong. \\
\textbf{Push It:} The curse also causes the target to suffer \textbf{1 Fatigue} as they struggle against the hostile woods. Mark \textbf{1 SB (Diamonds)} as the forest’s malice becomes visible. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the World’s Wound (High, 13 XP)} \hfill \\
\emph{Extended; Zone; Yes.} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[AREA]}, \texttt{[NATURE]}, \texttt{[TERRAIN]} \\
\textbf{Materials:} A sacred grove‑site, a circle of stones, an offering of blood or seed. \\
\textbf{Effect:} Heal a blighted place or quicken a damaged ecosystem. Clear taint, restore fertility, awaken dormant forces. Gain \textbf{+2 dice} to related \textit{Survival}/\textit{Nature} actions in the region. Start an \textbf{Ecosystem Restoration [8]} timer. \\
\textbf{Push It:} The cure spreads to adjacent areas but unbalances another; mark \textbf{2 SB (Diamonds)}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Cycle’s Crown (High, 14 XP)} \hfill \\
\emph{Scene; Self/Zone; Yes.} \\
\textbf{Tags:} \texttt{[TRANSFORM]}, \texttt{[NATURE]}, \texttt{[BOOST]}, \texttt{[PROTECT]} \\
\textbf{Materials:} A circlet of woven branches, antlers, and seasonal tokens. \\
\textbf{Effect:} Become an avatar of the cycle. Choose two:
\begin{itemize}
  \item \textbf{Seasonal Mastery:} \textbf{+2 dice} to all actions keyed to the current season (spring: healing and growth; summer: strength and harvest; autumn: preservation and preparation; winter: endurance and hiding).
  \item \textbf{Verdant Command:} Command plant life in \textit{Near} automatically.
  \item \textbf{Cycle’s Healing:} Remove one Condition or downgrade Harm by one level.
  \item \textbf{Growth’s Protection:} Allies gain \textbf{+1 Armor} from natural cover.
\end{itemize}
Begin a \textbf{Seasonal Avatar [6]} timer. \\
\textbf{Push It:} Extend the blessings to all present; mark \textbf{+2 Obligation} and gain \textbf{Seasonal Attunement} (see Corruption). \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 8 segments.

\subsubsection*{Cantors \& Cults of Grimmir}

\textbf{The Cantor’s Song.} A Cantor of Grimmir sings the \textit{drone of green things} – a low, humming chant that mimics the vibration of leaves in wind or the deep thrum of growing roots. Their voice is used to coax reluctant seeds, to calm wounded animals, and to mark the turning of the seasons in rituals that entire villages attend. A Grimmir Cantor does not seek an audience; they sing to the land itself, and the land answers in fruit, rain, or silence.

\textbf{The Cult of the Root‑Moot.} Grimmir’s cults are neither secret nor organised – a “moot” may be a single hermit or a scattered network of rangers who meet only at the solstices. The most ancient of these, the \textit{Root‑Moot}, carves masks from living moss and wears them during the rite of the Turning Wheel. Each member plants a tooth (their own or a beast’s) into the root‑ball of a sacred oak; if the tooth sprouts a sapling by spring, the planter is considered “heard” by the Old Man. Their creed is carved into the oldest oak in each grove: \textit{“Take only what the season gives. Give back what the season will need.”} Outsiders are welcome at the solstice fires, but they must work – splitting wood, digging a drainage trench, or feeding the fire with a personal memory (which the flames accept with a green flash).

\subsubsection*{Witchcraft of Grimmir (Lay / Threshold)}

A witch who walks with Grimmir is a \textit{hedgewright} or \textit{path‑keeper} – a crafter of living boundaries. Their work may be invisible: weaving a branch to guide deer away from a village, carving a marker that tells the frost when to end, or building a root bridge that only becomes passable at low tide. This craft can look like “infrastructure” – an Aeler True‑Mason might carve a stone that remembers the boundary between a mine and a sacred spring, or a Linn shipwright might season a keel with a prayer for the right wind. But the witch’s tools are never hammers; they are twine, moss, and the quiet patience of waiting for the land to answer.

\textbf{The Witch’s Tool: The Knotted Staff.} A staff of petrified wood, cold to the touch but warm when it touches living matter. Once per session, when used to bless a boundary marker (a post, a stone, a tree) it inscribes a minor \texttt{[SEASONAL]} tag into the threshold. The marker will “remember” the season of its carving for one year, granting \textbf{+1 die} to rolls that honour that season’s work.

\textbf{A Hedge Gift: Knot of the Turning Year (2 XP).} Once per scene, you may tie a simple knot (in a cord, a strand of hair, or a blade of grass) and speak a phrase appropriate to the current season. The knot subtly influences the weather or local plant growth for the next hour – a gust of wind, a sudden rain, or a patch of ivy that grows just enough to provide cover. The GM adjudicates effect.

\subsubsection*{Grimmir’s Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Seasonal Sense:} +1 die to predict weather, seasonal shifts, or natural events. & \textbf{Cyclical Mood:} –1 die to actions that defy the season (haste in winter, patience in spring). \\
2 & \textbf{Verdant Whisper:} Once/scene, gain +2 dice when communing with nature. & \textbf{Plant Speech:} Under stress, you address plants before people; –1 die to civilised social rolls. \\
3 & \textbf{Growth’s Touch:} Once/session, accelerate natural healing or growth. & \textbf{Seasonal Shift:} Demeanour changes with the season; allies note unsettling variance. \\
4 & \textbf{Boundary Keeper:} Once/scene, sense violations of wild places at \textit{Far}. & \textbf{Wild Intolerance:} –1 die in dense urban/cultivated zones; physical discomfort in worked stone. \\
5 & \textbf{Cycle’s Memory:} Once/session, perfectly recall a natural event witnessed here. & \textbf{Ecosystem Dependency:} Without daily time in nature, suffer \textbf{Fatigue 1}. \\
6+ & \textbf{Avatar of Seasons:} Once/session, embody a season fully (+3 dice to seasonal actions). & \textbf{Nature’s Claim:} Mark \textbf{+2 Obligation}; risk becoming bound to a specific biome. You may no longer eat cultivated food – only what you hunt, gather, or grow yourself. \\
\end{tblr}

\subsection*{Playstyle Notes}
Grimmir rewards those who approach the wild with fear and respect, not dominion. His followers are not masters of nature but participants in its rituals – they know when to give blood, when to leave an offering, and when to simply bow and walk away. Intrusions often take the form of seasonal demands: a grove that must be fed before the first frost, a creature that must be hunted before it hunts you, or a vision of a ritual that must be performed at the next solstice. Refusal does not invite wrath – merely the withdrawal of the wild’s protection, leaving the follower exposed to the very fears they once understood. Ideal for druids, rangers, and characters who understand that the forest is not a resource but a creditor.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Seasonal Sacrifice:} the old ways of blood and offering
  \item \textbf{Folk Horror:} the terror of the unseen and the ritual that must be performed
  \item \textbf{Primal Mystery:} knowledge hidden in root and stone, omens and portents
  \item \textbf{Growth from Death:} decay as the engine of new life
  \item \textbf{Boundary Keeping:} guarding the threshold between the civilised and the wild
\end{itemize}