% ============================================================================
% PATRON ENTRY: IKASHA, SHE WHO SLEEPS (LATENT POTENTIAL & SHADOW)
% ============================================================================

\subsection{Ikasha, She Who Sleeps (Latent Potential \& Shadow)}

\subsubsection*{Lore}
Ikasha is the hush between footfalls, the patience of dark water, the black-feathered watcher at every threshold. In stillness she gathers what might be, at crossroads she whispers of what may yet come. Ravens circle her, bearing secrets between worlds. Her followers learn to move unseen and speak unremembered, becoming shadows that slip between what is and what could be.

Ikasha does not act; she waits. She does not command; she suggests. Her power is not in force but in possibility -- the hidden door, the forgotten name, the secret that, if spoken, would change everything. Those who serve her become keepers of thresholds, listeners at the edge of sleep, and witnesses to the shape of things not yet born. She is the patron of spies, dreamers, outcasts, and those who know that the darkest shadow is also the safest cloak.

Among the peoples of Fate's Edge, Ikasha is called by many names. In the Mistlands, she is the \textit{Bell-Shrouded}, whose unknotted rope holds the fog at bay. In Aeler, she is the \textit{Deep Listener}, to whom miners whisper before descending into new shafts. In Zakov, she is the \textit{Tide Watcher}, a smuggler's patron who hides cargo from the Salt Prince's collectors. Wherever a threshold is crossed and a secret is kept, Ikasha watches.

\begin{quote}
``Blow out the candle. If the room listens back, ask softly. At the next crossroads, the raven waits -- and the shadow remembers your passing.''
\end{quote}

\subsubsection*{Domain Focus}
\begin{itemize}
  \item \textbf{Latent Potential:} the power of what is not yet, the seed of action, the whisper before the scream
  \item \textbf{Shadow as Shelter:} concealment, secrets, and the spaces between notice
  \item \textbf{Thresholds of Change:} doors, crossroads, dreams, and the boundaries of sleep
  \item \textbf{Raven's Witness:} omens, messages, and the weight of secrets shared
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}

A Runekeeper sworn to Ikasha does not bind a familiar of sunlight and noise. Their \textbf{Thiasos} is a small creature of shadow and stillness: a black cat with silver eyes that sees in perfect darkness, a gray mouse that nests in the walls of confessionals, a small owl whose feathers drink the light, a pale moth that appears only when a secret is spoken, or a raven chick with one white feather that knows the way to every crossroads. This creature requires no food, but it must spend at least an hour each day in absolute darkness (a closed drawer, a shuttered room). If neglected for a week, it becomes \textit{Compromised} (it grows restless and fidgets in light, imposing --1 die on all Stealth rolls for its bearer until appeased with a secret whispered in the dark).

The \textbf{Codex} of Ikasha is never a book read by lamplight. It may be a sheaf of blank pages that fill with whispered secrets only when the reader sleeps, a set of black silk ribbons knotted in patterns that change with the moon, or a slate that shows fading images of thresholds crossed. Because Ikasha's truths are written in the unwitnessed, the Codex requires upkeep: the Runekeeper must spend an hour each week in a dark room, letting the Codex rest in shadow, or the Rites stored there become faint (treat as Compromised until re-darkened).

\subsubsection*{Patron's Gift (Imbuement)}

Once per scene as an action, a Runekeeper of Ikasha may touch a held item (weapon, tool, or garment) and whisper a word of unlight. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Stealth} or \textit{Deception} rolls, and the first time each scene the bearer would be spotted by a mundane observer, they may reroll the opposed check (the shadow bends around them). After the scene, the user marks \textbf{+1 Obligation} to Ikasha.

\subsubsection*{Rites of Ikasha}

\paragraph{Rite of the Unlit Candle} (Basic, 6 XP) \hfill \\
\emph{Action; Self; Yes (Stealth).} \\
\textbf{Tags:} \texttt{[STEALTH]}, \texttt{[DECEPTION]}, \texttt{[SHADOW]} \\
\textbf{Materials:} A candle snuffed with a whisper. \\
\textbf{Effect:} You step into the space between notices. Gain \textbf{+1 die} to a single \textit{Stealth} or \textit{Subterfuge} roll, and if you succeed, you may also declare one minor detail ``forgotten'' by any observer (e.g., ``they don't remember my face,'' ``they left the key in the lock''). \\
\textbf{Push It:} Extend the effect to one ally within Near, but your shadow lingers independently for a moment -- the GM gains \textbf{1 SB (Diamonds)} as evidence remains. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Instant.

\paragraph{Rite of the Crossroads Raven} (Basic, 5 XP) \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[OMEN]}, \texttt{[NAVIGATION]}, \texttt{[SECRET]} \\
\textbf{Materials:} Three black feathers scattered at a crossing. \\
\textbf{Effect:} A raven appears, visible only to you and your allies. It offers an omen: you may reroll one failed navigation, pursuit, or escape roll this scene, or force one enemy to hesitate (lose their next minor action). \\
\textbf{Push It:} The raven whispers a true secret about the current situation (the GM must answer one yes/no question truthfully). In return, it demands a secret from you -- the GM gains \textbf{1 SB (Hearts)} to spend later as a complication involving that secret. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Umbral Reservoir} (Advanced, 8 XP) \hfill \\
\emph{Action; Self or Ally; No.} \\
\textbf{Tags:} \texttt{[SHADOW]}, \texttt{[RESTORATION]}, \texttt{[BOOST]} \\
\textbf{Materials:} Water gathered on a moonless night, held in a black vessel. \\
\textbf{Effect:} Draw on deep reserves of patience and shadow. Choose one:
\begin{itemize}
  \item Gain \textbf{+2 dice} on your next \textit{Stealth}, \textit{Deception}, or \textit{Resolve} roll.
  \item Clear \textbf{1 Fatigue} from yourself or a touched ally.
  \item Gain one free success on an attempt to escape restraint or pursuit (no roll needed).
\end{itemize}
\textbf{Push It:} Apply two benefits instead of one, but you mark \textbf{+1 Obligation} and the GM gains \textbf{1 SB (Clubs)} as the shadow's debt comes due. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Instant.

\paragraph{Rite of the Unlatched Secret} (Advanced, 9 XP) \hfill \\
\emph{Instant; Near; Yes.} \\
\textbf{Tags:} \texttt{[SECRET]}, \texttt{[TRUTH]}, \texttt{[UNSEAL]}, \texttt{[READ]} \\
\textbf{Materials:} A key that has never been used, bent in a circle, or a lock of hair / personal token / name written on black paper. \\
\textbf{Effect:} Choose one of the following:
\begin{itemize}
  \item \textbf{Unlatched Door:} Persuade a mundane threshold to open. One lock, seal, or barrier (door, chain, shutter, even a pursuer's grip) opens or loosens as if never secured. Does not work on magical wards.
  \item \textbf{Secret Keeper:} Compel a truthful answer to one direct question about a secret the target knows. They resist with Resolve (DV 4). On failure, they answer truthfully (cryptically permitted). On success, they suffer \textbf{1 Fatigue} from concealment.
\end{itemize}
\textbf{Push It:} If you chose \textit{Unlatched Door}, you may also silence any noise the opening makes, but the threshold ``remembers'' you -- the GM gains \textbf{1 SB (Spades)} for a later complication. If you chose \textit{Secret Keeper}, you also learn a hidden emotion or related secret (GM's choice), but the target learns one of your secrets -- the GM gains \textbf{1 SB (Hearts)} to reveal it later. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Instant.

\paragraph{Rite of the Umbral Dream} (Master, 14 XP) \hfill \\
\emph{Extended; Self; No.} \\
\textbf{Tags:} \texttt{[SHADOW]}, \texttt{[DREAM]}, \texttt{[FATE]}, \texttt{[TRANSPORT]}, \texttt{[INVISIBLE]} \\
\textbf{Materials:} Stand in absolute darkness or at a deserted crossroads at midnight; or burn a vivid dream written down. \\
\textbf{Effect:} You become a living shadow or reach into latent potential. Choose one of the following:
\begin{itemize}
  \item \textbf{Shadow Form:} For the scene, you are intangible to mundane harm, can pass through small gaps and thresholds, gain \textbf{+2 dice} to all \textit{Stealth} rolls, and automatically succeed on one escape attempt. You cannot manipulate objects normally, and you cannot carry items through shadows or bring objects with you when you become intangible. Any items you are carrying fall to the ground (or to the surface you were on) when you assume this form; you may retrieve them after the effect ends.
  \item \textbf{Whisper of Possibility:} Once before the next dawn, reroll any single roll after seeing the result (before outcome is declared). The GM gains \textbf{1 SB} as fate wavers.
  \item \textbf{Hidden Path:} You and allies bypass one major obstacle (guarded gate, collapsed bridge, social dead end) as if a secret way had always been there -- the GM gains \textbf{1 SB} as the world adjusts.
  \item \textbf{Unspoken Name:} Learn the true name of one creature or place (GM discretion). That name can be used for future rites or bargains.
\end{itemize}
\textbf{Push It:} Apply two benefits instead of one. If you chose \textit{Shadow Form}, you may also interact once with a locked or warded object while intangible, but your next defense roll is \textit{Desperate}. In all cases, the GM gains \textbf{2 SB} (the dream's echoes attract attention). Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Scene (or as specified).

\subsubsection*{Cantors \& Cults of Ikasha}

\textbf{The Cantor's Song.} A Cantor of Ikasha sings in \textit{whispers} -- melodies that are felt rather than heard, harmonies that slip beneath the threshold of attention. Their voice is used to hide a crowd's escape, to make a guard forget why they entered a room, or to pass a message through enemy lines without a single word being spoken aloud. Ikasha's Cantors are rarely recognized; they are the ones who hum absently while others plan rebellion.

\textbf{The Cult of the Unlit Candle.} Ikasha's cults are shadowy by necessity, meeting in basements, abandoned bell-towers, and fog-bound crossroads. They whisper freedom to the oppressed -- escaped slaves, hunted heretics, debtors fleeing their creditors. A ritual of the Unlit Candle involves each member extinguishing a flame and speaking a secret wish for escape. The candles are never relit; the wish is carried away on ravens' wings. Their creed is never written, only passed from shadow to shadow: \textit{``The dark does not confine -- it conceals the road that the light would forbid.''} Outsiders who stumble into a meeting are offered a black feather and a choice: forget, or follow.

\subsubsection*{Witchcraft of Ikasha (Lay / Shadow)}

A witch who walks with Ikasha is a \textit{threshold-keeper} or \textit{shadow-weaver} -- one who knows that a door left unlatched is not an invitation to theft, but to liberation. Their magic works in the gaps: the crack under a door, the moment between heartbeats, the pause between a guard's patrol. This craft can be invisible -- a bent lock, a misplaced key, a shadow that moves of its own accord -- or as grand as erasing a name from a pursuer's memory.

\textbf{The Witch's Tool: The Raven's Quill.} A quill carved from a raven's primary feather, dipped not in ink but in the absence of light. Once per session, when used to write a name on a threshold (a doorframe, a gatepost, a windowsill), that name becomes ``forgotten'' for a scene: the person whose name is written cannot be perceived by those who cross that threshold until they succeed on a Resolve test (DV 3). The quill can also erase a single word from a written document (it becomes a smudge of shadow).

\textbf{A Hedge Gift: Unlatched (Basic, 2 XP).} Once per scene, you may declare a mundane lock, latch, or binding to be ``unlatched'' -- it opens freely (or fails to close) as if it had never been secured. This works on doors, chains, shutters, and even a pursuer's grip, but not on magical wards.

\subsubsection*{Corruption of Ikasha}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Shadow's Whisper:} +1 die to \textit{Deception} when lying in darkness or through intermediaries. & \textbf{Truth Echoes:} Occasionally speak in riddles or half-truths without realizing it. \\
2 & \textbf{Umbral Sight:} Once per scene, gain +2 dice to \textit{Notice} hidden threats or ambushes in dim light. & \textbf{Light Sensitivity:} --1 die to rolls requiring keen vision in bright conditions. \\
3 & \textbf{Raven's Memory:} Never forget a secret told to you; recall any whispered conversation perfectly. & \textbf{Secret Burden:} Must keep one troubling secret per Tier; suffer \textbf{1 Fatigue} when actively trying to forget. \\
4 & \textbf{Threshold Walker:} Once per session, move through one locked door, sealed letter, or magical barrier without triggering wards. & \textbf{Crossroads Bound:} Must pause at literal or metaphorical crossroads to ``listen'' before major decisions. \\
5 & \textbf{Shadow Debt:} Once per session, call in a favor from someone who owes you a secret -- no questions asked. & \textbf{Obligation Web:} Every secret you learn creates a subtle tie to its keeper; mark \textbf{1 SB (Diamonds)} when ignoring these connections. \\
6+ & \textbf{Umbral Ascendancy:} Once per session, become completely undetectable to all non-magical senses for one scene. & \textbf{Reality Thinning:} The boundary between what is and what could be grows thin -- the GM may introduce subtle reality alterations around you. \\
\end{tblr}

\subsection*{Playstyle Notes}
Ikasha rewards patience, subtlety, and the willingness to let things unfold. Her followers are not agents of action but of possibility -- they open doors, plant secrets, and wait for the harvest. Intrusions often take the form of dreams showing a hidden danger, a raven appearing with a cryptic message, or a secret you thought buried resurfacing at the worst moment. Refusing an intrusion does not bring wrath -- merely the withdrawal of Ikasha's favor: doors stay locked, shadows become just shadows, and the road ahead grows a little darker. Ideal for rogues, spies, and characters who understand that the greatest power is often the one that remains unseen.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Latent Potential:} power as possibility, not force
  \item \textbf{Shadow as Shelter:} concealment, secrets, and the spaces between notice
  \item \textbf{Thresholds of Change:} doors, crossroads, dreams, and the boundaries of sleep
  \item \textbf{Raven's Witness:} omens, messages, and the weight of secrets shared
\end{itemize}

% --- Monastic Tradition: Shadow Monk (Ikasha) ---
\begin{tcolorbox}[colback=black!5,colframe=black!70!white,title=\textbf{Shadow Monk (Ikasha)},breakable]
  \emph{Blow out the candle. If the room listens back, ask softly.}

  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Wits 3+, Spirit 2+, Rite of the Unlit Candle or Cantor of Ikasha.

  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} Shadows leave no mark on ledgers. Monks of the Shadow Monk gain \textbf{+1 die} to resist any magical debt, geas, or contract that would compel them to be seen, remembered, or witnessed. If they succeed, the creditor's records show only a blank space where their name should be -- a gap the Hollow may one day fill.

  \vspace{2mm}
  \noindent\textbf{Techniques (purchased as Talents):}

  \begin{tblr}{
    colspec = {X[0.7,l] X[1.8,l] X[1.1,l] X[1.4,l]},
    rowhead = 1,
    row{odd}={bg=gray!10},
    row{1}={bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Shadow Step} (Basic, 6 XP) & Move from one patch of shadow to another within Near range without crossing the space between. You must be able to see both shadows. & 1 Fatigue & Wits 3 \\
  \textbf{Unlit Candle} (Basic, 6 XP) & Step into the space between notices. Gain +1 die to a single Stealth or Subterfuge roll, and if you succeed, declare one minor detail ``forgotten'' by any observer. & 1 Boon & Spirit 2 \\
  \textbf{Crossroads Raven} (Basic, 5 XP) & A raven appears, visible only to you and your allies. Reroll one failed navigation, pursuit, or escape roll this scene, or force one enemy to lose their next minor action. & 1 Fatigue & Wits 2 \\
  \textbf{Unlatched Secret} (Advanced, 9 XP) & Persuade a mundane threshold to open (door, lock, even a pursuer's grip). Does not work on magical wards. & 1 Boon & Spirit 3 \\
  \textbf{Umbral Reservoir} (Advanced, 8 XP) & Draw on deep reserves of patience and shadow. Clear 1 Fatigue from yourself or a touched ally, or gain +2 dice on your next Stealth, Deception, or Resolve roll. & 1 Fatigue & Tier II \\
  \textbf{Hidden Path} (Advanced, 10 XP) & You and allies bypass one major obstacle (guarded gate, collapsed bridge, social dead end) as if a secret way had always been there -- the GM gains 1 SB as the world adjusts. & 1 Boon & Tier III \\
  \end{tblr}

  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 14 XP):}
  \textit{Shadow Form.} Once per session, become intangible to mundane harm for one scene: pass through small gaps, gain +2 dice to all Stealth rolls, and automatically succeed on one escape attempt. You cannot manipulate objects normally. After the scene, you collapse with Fatigue 2 and gain a permanent \textbf{Debt Scar} (your shadow is faint, as if you borrowed it from someone else). You may ignore one forced Obligation or debt call -- the creditor sees only a silhouette that dissolves when touched.
\end{tcolorbox}