% --- Patron: Inaea, Angel of the Spider (Webs, Fate \& Patient Predation) ---
\subsection{Inaea, Angel of the Spider (Webs, Fate \& Patient Predation)}

\textit{Lore.} Inaea sits in patient silence at the center of the unseen lattice. Where others see isolation, she sees threads---debt, affection, rivalry, loyalty---all binding mortals to one another. She is comforter and conspirator alike: the hearth-mother who eases grief with a warm hand, and the shadow-spinner who knots fates so they cannot slip free. Those who serve her learn to pull strands, tying allies closer and ensnaring foes until mercy or trap is inevitable.

The spider does not rush. It weaves, waits, and feels the tremor of every caught fly. So too does Inaea. Her mercy is never without cost; her kindness is a thread that, once accepted, can be pulled taut at a moment of her follower's choosing. To serve her is to understand that every bond is a leash, every promise a snare, and every act of compassion a line cast into the dark.

Among the peoples of Fate's Edge, Inaea is the patron of hearth-keepers, matchmakers, and those who bind communities. In Aelaerem, she is the \textit{Kettle-Spinner}, invoked by grandmothers to keep families together. In Silkstrand, she is the \textit{Dyed Web}, honored by dyers' guilds who leave a single strand of silk outside their windows on the first of spring. Wherever threads are tied---in marriage, in treaty, in the quiet forgiveness of a debt---Inaea's influence is felt.

\begin{quote}
``The thread that comforts is the same that binds. The knot that saves is also the snare. What you accept from my hand, you give me leave to tighten.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Webs of Obligation:} debts, promises, favors, and the ties that bind
  \item \textbf{Patient Predation:} setting traps, waiting for the moment to strike
  \item \textbf{Unseen Connections:} sensing hidden relationships, leverage, and weakness
  \item \textbf{Mercy as Control:} healing and aid that leave lasting threads
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Inaea does not bind a familiar of claw and fang. Their \textbf{Thiasos} is a small creature of threads and patience: a garden spider that weaves its web in the shape of the keeper's name, a tiny weaver bird that knots grass into intricate patterns, a small silk moth that leaves a trail of thread wherever it flies, or a gray field mouse that nests in bundles of twine. This creature requires no food, but it must be given a fresh length of thread, ribbon, or cloth to work with each day. If neglected for a week, it becomes \textit{Compromised} (it tangles cords and knots laces at inopportune moments, imposing ---1 die on all fine manipulation rolls for its bearer until offered a new thread).

The \textbf{Codex} of Inaea is never a book of ink and paper. It may be a frame loom with a half-finished weaving that records Rites as colored threads, a bundle of knotted cords that must be read by touch, or a wooden spindle wrapped with silk strands that sing when plucked. Because Inaea's truths are woven, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week adding a new row to the weaving (or untying and retying a knot), or the patterns become tangled (treat as Compromised until re-woven).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Inaea may touch a held item (weapon, tool, or garment) and whisper a word of binding. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Sway} or \textit{Insight} rolls when dealing with promises, debts, or relationships, and the first time each scene the bearer would be betrayed by an ally under a promise, they may sense it one exchange in advance (the thread pulls taut). After the scene, the user marks \textbf{+1 Obligation} as a thread of connection tightens around them.

\subsubsection*{Rites of Inaea}

\paragraph{Hearth-Thread Knot (Low, 5 XP)} \hfill \\
\emph{Scene; Near; Yes (link once).} \\
\textbf{Tags:} \texttt{[BOND]}, \texttt{[PROTECT]}, \texttt{[HEAL]} \\
\textbf{Materials:} A thread tied around a wrist or shared cup. \\
\textbf{Effect:} Bind yourself and one willing target (ally or neutral) with a sympathetic thread. Both gain \textbf{+1 die} to Aid each other this scene. Additionally, if either takes Harm, the other may reduce it by 1 (to a minimum of 0) by suffering that same Harm instead. This transfer is voluntary for an ally; for an unwilling target, you must succeed on a Resolve test (DV 3) to force the link. \\
\textbf{Push It:} Extend to two allies; mark \textbf{1 SB (Hearts)} as your own lifeline strains. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\emph{XP Cost:} 5 XP (Basic Technique)

\paragraph{Snaring Filament (Low, 5 XP)} \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[TRAP]}, \texttt{[AREA]}, \texttt{[BIND]} \\
\textbf{Materials:} A knotted cord or woven scrap. \\
\textbf{Effect:} Lay an invisible snare in a lane, doorway, or narrow passage. The first hostile creature to cross suffers \textbf{--2 dice} on their next action and must succeed on a Resolve test (DV 2) or lose their movement for that turn (tangled). The snare lasts for the scene or until triggered. \\
\textbf{Push It:} Lay a second snare elsewhere, but allies risk entanglement if they forget its location; mark \textbf{1 SB (Spades)} as the web spreads thin. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\emph{XP Cost:} 5 XP (Basic Technique)

\paragraph{Strand of Inevitability (Standard, 8 XP)} \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[BOND]}, \texttt{[FATE]}, \texttt{[MANIPULATION]} \\
\textbf{Materials:} Three colored threads (white, black, red) braided together. \\
\textbf{Effect:} Link two actors (which can include yourself) with a visible or invisible strand. For the rest of the scene, whenever one acts, the other is dragged into consequence. You may trigger this linkage once to either:
\begin{itemize}
  \item Force \textbf{1 SB (Hearts)} on the linked target after they perform a significant action.
  \item Gain \textbf{+2 dice} on a \textit{Setup}, \textit{Predict}, or \textit{Manipulate} roll that exploits the link (e.g., ``you cannot betray me without hurting your ally'').
\end{itemize}
The link is one-way or two-way as you choose. \\
\textbf{Push It:} Invert the direction of the link once after it is set (e.g., make the pursuer stumble instead of the pursued). Mark \textbf{1 SB (Diamonds)} as the weave resists. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\emph{XP Cost:} 8 XP (Advanced Technique)

\paragraph{Weaver's Glance (Standard, 7 XP)} \hfill \\
\emph{Instant; Self; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[SENSE]},  \\
\textbf{Materials:} Observe a spider finishing a web's radial line (or a convincing imitation). \\
\textbf{Effect:} Ask the GM one question about a hidden tie, debt, or leverage between two people, or between a person and a faction. The GM must answer truthfully. Your next \textit{Manipulate}, \textit{Pressure}, or \textit{Intimidate} action exploiting that knowledge gains \textbf{+1 Effect}. \\
\textbf{Push It:} Reveal a secret tie that was not previously declared in the fiction (the GM must invent a plausible connection). Mark \textbf{Exposure +1} as your own entanglements surface. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\emph{XP Cost:} 7 XP (Basic/Advanced borderline; treat as 7 XP)

\paragraph{Spinner's Debt (Standard, 9 XP)} \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[DEBT]}, \texttt{[BOND]}, \texttt{[COMMAND]} \\
\textbf{Materials:} A silver coin and a strand of your own hair, knotted together. \\
\textbf{Effect:} Offer a willing creature a boon: they may immediately clear \textbf{1 Fatigue} or gain \textbf{+1 die} on their next roll. In return, they accept a \textbf{Promise Timer [4]} named ``Debt to the Web.'' You may advance this timer by 1 segment at any time as a free action (once per scene). When the timer fills, you may demand a single service from them --- not impossible, but significant (e.g., ``speak for me at court,'' ``reveal a secret,'' ``take a blow meant for me''). Refusal causes them to suffer \textbf{Harm 1 (Stress)} and the timer resets. \\
\textbf{Push It:} Extend the offer to a group (up to three targets). Each who accepts gives you a separate Promise Timer. Mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\emph{XP Cost:} 9 XP (Advanced Technique)

\paragraph{Binding Knot (High, 11 XP)} \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[BOND]}, \texttt{[FATE]}, \texttt{[COMMAND]} \\
\textbf{Materials:} A silk loop tied between two hands, then cut or knotted with a spoken vow. \\
\textbf{Effect:} Bind two parties (individuals or small groups) to a mutual oath. Breaking the vow forces \textbf{2 SB} on the violator and marks them with a subtle tell (a visible thread, a nervous twitch, a lingering scent) that persists for one week. \\
\textbf{Push It:} Widen the binding to up to four parties, each of whom may define one loophole. Exploiting a loophole causes \textbf{1 SB (Clubs)} as the web tightens around the exploiter. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments. \\
\emph{XP Cost:} 11 XP (Advanced/Master borderline)

\paragraph{Merciful Severing (High, 13 XP)} \hfill \\
\emph{Scene; Touch; No.} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[DISPEL]}, \texttt{[MERCY]}, \texttt{[DEBT]} \\
\textbf{Materials:} A white thread burned to ash, spoken with a name. \\
\textbf{Effect:} Sever one harmful tie --- a curse, a toxic obsession, a coercive bond, or a supernatural debt. Remove one Condition or \textbf{1 SB} tied to it. This is an act of genuine mercy; the target may be unwilling, but the thread can still be cut (Resolve test DV 4 to resist). \\
\textbf{Push It:} Sever a tie without the target's consent or knowledge; they will not know who cut them free, but you mark \textbf{+2 Obligation} as the web recoils against you, and the severed strand may attach to you instead (GM's discretion). \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments. \\
\emph{XP Cost:} 13 XP (Master Technique)

\paragraph{Web of Slumber (High, 12 XP)} \hfill \\
\emph{Extended; Zone; No.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[SLEEP]}, \texttt{[BIND]}, \texttt{[FEAR]} \\
\textbf{Materials:} A complete spider's web, stolen at dawn, and a lullaby hummed into it. \\
\textbf{Effect:} Weave an area of drowsy stillness. For the next hour, any creature entering the zone must succeed on a Resolve test (DV 4) or fall into a light trance, losing their next action and suffering \textbf{--1 die} to all perception rolls while inside. Allies you touch beforehand are immune. The web is invisible until triggered. \\
\textbf{Push It:} The trance becomes deeper --- sleeping creatures cannot be woken by non-magical means for the duration. Mark \textbf{2 SB (Spades/Hearts)} as the web draws hungry attention. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments. \\
\emph{XP Cost:} 12 XP (Master Technique)

\subsubsection*{Cantors \& Cults of Inaea}

\textbf{The Cantor's Song.} A Cantor of Inaea sings in \textit{knots} --- melodies that loop back on themselves, harmonies that tie two voices into one. Their voice is used to heal quarrels, to bind a community together after a tragedy, or to weave a subtle compulsion into a seemingly innocent lullaby. Inaea's Cantors are beloved at weddings, feared at treaty signings, and always watched by those who understand that a pretty tune can be a leash.

\textbf{The Cult of the Woven Circle.} Inaea's cults gather in domestic spaces --- kitchens, weaving sheds, longhouses --- and resemble extended families more than secret societies. A ritual of the Woven Circle involves each member tying a colored thread around their wrist, then passing the ends to their neighbors. The resulting web is then tightened, and a communal promise spoken. The cult's creed is embroidered onto a single cushion that passes from host to host: \textit{``No thread stands alone. The weakest knot still holds the strongest weight.''} Outsiders are welcome if they bring a thread and a truth. Betrayal of the circle is repaid by being unraveled --- socially, financially, or worse.

\subsubsection*{Witchcraft of Inaea (Craft / Threshold)}

A witch who walks with Inaea is a \textit{web-weaver} or \textit{knot-smith} --- one who understands that every threshold (a door, a marriage, a treaty) is held together by invisible threads. Their magic works with domestic tools: a spinning wheel that twists fate, a loom that weaves obligations into cloth, or a pair of shears that cuts a single strand to end a feud. This craft can be as humble as tying a red thread around a child's wrist for protection, or as grand as weaving a tapestry that records an entire bloodline's debts.

\textbf{The Witch's Tool: The True Loom.} A small, portable loom --- no larger than a book --- woven from hawthorn wood and strung with thread dyed in three colors (red for blood, white for truth, black for secret). Once per session, the witch may weave a single knot into a new or existing thread; that knot becomes a \emph{promise timer} [4]. When the timer fills, the promise is either fulfilled or broken, as agreed at the time of weaving.

\textbf{A Hedge Gift: Bind the Threshold (2 XP).} Once per scene, the witch may tie a simple knot in a cord, ribbon, or strand of hair and place it on a threshold (a doorstep, a gate, a bridge). For the next hour, anyone crossing that threshold must succeed on a Resolve test (DV 2) if they intend to break an oath, tell a lie, or harm someone inside. The knot visibly tightens if a violation is attempted. \\
\emph{XP Cost:} 2 XP (Minor hedge gift; not a full technique)

\subsubsection*{Inaea's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1.5cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Hearth-Blessed:} +1 die to \textit{Comfort} or \textit{Aid} actions. & Must offer solace when asked, or mark \textbf{1 SB (Hearts)}. \\
2 & \textbf{Web-Sense:} Once/scene, sense hidden ties between two people (debt, rivalry, affection). & \textbf{Empathic Overload:} suffer \textbf{1 Fatigue} when immersed in strong conflict. \\
3 & \textbf{Binding Presence:} Once/session, compel a promise unless target passes Resolve DV 4. & \textbf{Over-Attachment:} --1 die when abandoning those you've aided or bound. \\
4 & \textbf{Fate's Pattern:} +2 dice to predict alliances, betrayals, or bargains. & \textbf{Manipulative Urge:} mark \textbf{1 SB (Diamonds)} when acting selflessly without agenda. \\
5 & \textbf{Spider's Mercy:} Once/session, sever one harmful tie for yourself or another (as \textit{Merciful Severing}). & \textbf{Hunger for Threads:} suffer \textbf{1 Fatigue} if a session passes without weaving a new tie or extracting a promise. \\
6+ & \textbf{Weaver's Dominion:} Once/session, declare all ties in a scene bend toward you; gain +2 dice on social manipulation. & \textbf{Entanglement:} Mark \textbf{+2 Obligation}; those affected become subtly dependent on your will. \\
\end{tblr}

\subsection*{Playstyle Notes}
Inaea rewards those who see relationships as both comfort and leverage. Her followers are not merely kind; they are patient, observant, and willing to let a debt ripen before collecting. Intrusions often take the form of a sudden obligation --- a person you helped now demands help in return, a promise you thought forgotten tightens, or a web you spun catches someone you did not intend. Refusing an intrusion does not break the thread, but it frays it: the GM may mark a \textit{Tangled Web} timer that, when filled, causes all your active promises and bonds to tighten at once, often at the worst moment. Ideal for healers, diplomats, and characters who understand that mercy is a thread that can be pulled.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Patient Predation:} setting threads and waiting for the right moment
  \item \textbf{Webs of Obligation:} debts, promises, and favors as both comfort and leash
  \item \textbf{Unseen Connections:} sensing hidden relationships and leverage
  \item \textbf{Mercy as Control:} healing and aid that leave lasting threads
\end{itemize}