% ============================================================================
% patrons/index.tex — The Invoker's Catalogue
%
% This file is the master index for all patron entries. It is included by
% the main Invoker's Guide document. Each patron is stored in its own .tex file
% in the patrons/ subdirectory.
%
% The structure is built around the Triune taxonomy: three Orders of eleven
% Thrones each, plus a fourth category for Patrons beyond that architecture.
%
% All \input paths are relative to the patrons/ directory.
% ============================================================================

% --- PREAMBLE: THE INVOKER'S VIEW ---

\section*{The Invoker's Catalogue}
\addcontentsline{toc}{section}{The Invoker's Catalogue}

\begin{quote}
``I have carried a dozen Symbols and broken half of them. The weight does not diminish; it only redistributes. This is the Invoker's art: to know which Patron to call, when to call them, and what they will demand in return. The Triune is not a religion. It is a map.''
\end{quote}

\noindent
For the Invoker, the Patrons are not gods to be worshipped but \emph{forces to be understood}. They are the great constants of the world—tides, storms, laws, hungers, and silences—each with its own nature, its own demands, and its own price. The Invoker does not kneel; they \emph{study}. They learn the shape of a Patron's attention, the weight of its Obligation, and the precise moment when a Symbol must be displayed, a Rite invoked, or a Seal cracked.

The following catalogue presents the Patrons according to the Triune taxonomy—a framework not of worship, but of \emph{comprehension}. The Triune divides the Patrons into three great Orders, each ruled by one of the three sisters who are not sisters, the three angels who are not angels:

\begin{itemize}[leftmargin=1.5em]
  \item \textbf{The Seraphic Order} — law, flame, and connection. The principles that bind and structure.
  \item \textbf{The Empyreal Order} — motion, desire, and performance. The forces that drive and transform.
  \item \textbf{The Chthonic Order} — depth, memory, and endings. The powers that dissolve and renew.
\end{itemize}

\noindent
Beyond these three lie the Patrons who stand outside the architecture—too singular, too wild, or too newly formed to be contained. They are not lesser; they are \emph{unclassifiable}. The wise Invoker knows when to reach for the known and when to reach for the strange.

\vspace{1em}

\noindent
\textbf{How to Read This Catalogue:}
\begin{itemize}[leftmargin=1.5em]
  \item Each Patron is listed with its \textbf{Order} (if any), \textbf{Throne number}, \textbf{domain}, and \textbf{aspect}.
  \item The \textbf{Invoker's Note} offers practical guidance on when and why to invoke that Patron.
  \item \textbf{Acts of Service} are provided as examples of how to clear Obligation—each Patron rewards resonance, not mere labour.
  \item \textbf{Rivalries} indicate which Patrons resist each other; carrying Symbols from rivals creates Cross-Resonance and narrative friction.
  \item \textbf{Thiasos} examples are given for Runekeepers; for the Invoker, these serve as models for the kind of attention each Patron prefers.
\end{itemize}

\newpage

% ============================================================================
% THE SERAPHIC ORDER — CHOIR OF LAW, FLAME, AND CONNECTION
% ============================================================================

\section{The Seraphic Order}
\label{sec:seraphic-order}

\begin{quote}
\textit{Ruled by Inaea, Angel of the Spider} \\
``In the web, all vows are caught.''
\end{quote}

The Seraphic Order embodies the principles of \textbf{structure, purity, and mercy}. Its Patrons govern covenants, justice, and the bonds that hold society together. They are the \emph{law} that enables trust, the \emph{flame} that purifies corruption, and the \emph{mercy} that tempers judgment.

For the Invoker, the Seraphic Patrons offer reliable, structured power. Their Rites are predictable, their Obligation costs moderate, and their Acts of Service straightforward. The price is \emph{accountability}: these Patrons remember what you have sworn, and they hold you to it.

\newpage

% --- THRONES I–XI: SERAPHIC ORDER ---
\input{patrons/inaea.tex}              % I — Inaea
\newpage
\input{patrons/oath-flame-light.tex}   % II — Oath of Flame & Light
\newpage
\input{patrons/mykkiel.tex}            % III — Mykkiel
\newpage
\input{patrons/sacred-geometry.tex}    % IV — Sacred Geometry
\newpage
\input{patrons/inquisitor-prime.tex}   % V — Inquisitor Prime
\newpage
\input{patrons/gallows-bell.tex}       % VI — Gallows Bell
\newpage
\input{patrons/varnek-karn.tex}        % VII — Varnek Karn
\newpage
\input{patrons/confessor.tex}          % VIII — Confessor Beneath the Bell
\newpage
\input{patrons/silent-choir.tex}       % IX — Silent Choir
\newpage
\input{patrons/witness.tex}            % X — The Witness
\newpage
\input{patrons/sealed-gate.tex}        % XI — Sealed Gate

% ============================================================================
% THE EMPYREAL ORDER — CHOIR OF MOTION, DESIRE, AND PERFORMANCE
% ============================================================================

\section{The Empyreal Order}
\label{sec:empyreal-order}

\begin{quote}
\textit{Ruled by Ikasha, She Who Sleeps} \\
``In shadow and silence, all things move.''
\end{quote}

The Empyreal Order embodies the principles of \textbf{fluidity, hunger, and experience}. Its Patrons govern desire, performance, and the motion that drives all change. They are the \emph{dance} that fills the void, the \emph{hunger} that seeks fulfilment, and the \emph{journey} that never ends.

For the Invoker, the Empyreal Patrons offer flexibility and passion. Their Rites are rewarding but unpredictable; their Obligation costs are moderate, but their demands are \emph{emotional}. The price is \emph{vulnerability}: these Patrons ask you to feel, to want, to risk.

\newpage

% --- THRONES I–XI: EMPYREAL ORDER ---
\input{patrons/ikasha.tex}             % I — Ikasha
\newpage
\input{patrons/raeyn.tex}              % II — Raéyn
\newpage
\input{patrons/thrysos.tex}            % III — Thrysos
\newpage
\input{patrons/mab.tex}                % IV — Mab
\newpage
\input{patrons/livaea.tex}             % V — Livaea
\newpage
\input{patrons/palinode.tex}           % VI — Palinode
\newpage
\input{patrons/aveh.tex}               % VII — Aveh
\newpage
\input{patrons/morag.tex}              % VIII — Morag the Hag
\newpage
\input{patrons/grimmir.tex}            % IX — Grimmir
\newpage
\input{patrons/maelstraeus.tex}        % X — Maelstraeus
\newpage
\input{patrons/traveler.tex}           % XI — The Traveler

% ============================================================================
% THE CHTHONIC ORDER — CHOIR OF DEPTH, MEMORY, AND ENDINGS
% ============================================================================

\section{The Chthonic Order}
\label{sec:chthonic-order}

\begin{quote}
\textit{Ruled by Isoka, Angel of Serpents} \\
``Every skin is a memory you no longer need.''
\end{quote}

The Chthonic Order embodies the principles of \textbf{dissolution, renewal, and silence}. Its Patrons govern decay, transformation, and the endings that make new beginnings possible. They are the \emph{serpent} that sheds its skin, the \emph{void} that consumes memory, and the \emph{silence} that follows the song.

For the Invoker, the Chthonic Patrons offer transformative and absolute power. Their Rites are potent but costly; their Obligation costs are high, and their Acts of Service are demanding. The price is \emph{transformation}: these Patrons ask you to change, to shed, to end.

\newpage

% --- THRONES I–XI: CHTHONIC ORDER ---
\input{patrons/isoka.tex}              % I — Isoka
\newpage
\input{patrons/malachai.tex}           % II — Malachai
\newpage
\input{patrons/carrion-king.tex}       % III — Carrion King
\newpage
\input{patrons/nidhoggr.tex}           % IV — Nidhoggr
\newpage
\input{patrons/khemesh.tex}            % V — Khemesh
\newpage
\input{patrons/clockwork-monad.tex}    % VI — Clockwork Monad
\newpage
\input{patrons/pale-shepherd.tex}      % VII — Pale Shepherd
\newpage
\input{patrons/vorthak.tex}            % VIII — Vorthak
\newpage
\input{patrons/ninth-rim.tex}          % IX — The Ninth
\newpage
\input{patrons/mor-iraath.tex}         % X — Mor'iraath
\newpage
\input{patrons/ninth-rim.tex}          % XI — Ninth Rim

% ============================================================================
% PATRONS BEYOND THE TRIUNE — WILD POWERS
% ============================================================================

\section{Patrons Beyond the Triune}
\label{sec:patrons-beyond}

Not all Patrons fit neatly into the Triune's three Orders. Some are older than the taxonomy itself; others are too singular, too wild, or too newly formed to be contained by the architecture of the Thirty-Three Thrones. They are not lesser—they are \emph{unclassifiable}. An Invoker who understands them holds power that the orthodox do not.

For the Invoker, these Patrons offer specialised, unconventional power. Their Rites are often unique, their Obligation costs vary widely, and their Acts of Service are tailored to their specific natures. The price is \emph{singularity}: these Patrons ask you to embrace their particular vision of the world.

\newpage

% --- PATRONS BEYOND THE TRIUNE ---
\input{patrons/breath-first-forge.tex} % The Breath of the First Forge — Creativity, Ingenuity & the Joy of Making
\newpage
\input{patrons/kuva.tex}               % Kuva — Sky, Storms, Witness
\newpage
\input{patrons/lucky-jack.tex}         % Lucky Jack — Luck & The Heist
\newpage
\input{patrons/lunara.tex}             % Lunara — Moon & Mystery
\newpage
\input{patrons/oya.tex}                % Oya — Wind of the Sahel, Liberation
\newpage
\input{patrons/riven_path.tex}         % Riven Path — Passion & Power
\newpage
\input{patrons/savage-heart.tex}       % Savage Heart — Primal Consumption
\newpage
\input{patrons/seal-breaker.tex}       % Seal-Breaker — Forgotten Names & Broken Seals
\newpage
\input{patrons/unbroken_way.tex}       % Unbroken Way — Balance & Control
\newpage
\input{patrons/xhak-thul.tex}          % Xhak'Thul — Storm, Shock & Awe

% ============================================================================
% INVOKER REFERENCE TABLES
% ============================================================================

\clearpage
\section{Invoker's Reference Tables}
\label{sec:invoker-tables}

\subsection{Patron's Gift — Thematic Skills and Bestowals}
\label{subsec:gift-table}

The following table summarises the \textbf{Patron's Gift} granted to Runekeepers and the \textbf{Borrowed Grace} effect for Invokers. The +1 die applies only when the fiction aligns with the Patron's domain.

\renewcommand{\arraystretch}{1.15}
\setlength{\LTpre}{0.5em}
\setlength{\LTpost}{0.5em}

\begin{longtable}{@{}p{4.2cm}p{3.2cm}p{7.2cm}@{}}
  \toprule
  \textbf{Patron} & \textbf{Thematic Skill (+1)} & \textbf{Gift / Lore Bestowal} \\
  \midrule
  \endfirsthead
  
  \toprule
  \textbf{Patron} & \textbf{Thematic Skill (+1)} & \textbf{Gift / Lore Bestowal} \\
  \midrule
  \endhead
  
  \midrule
  \multicolumn{3}{r}{\emph{Continued on next page}} \\
  \bottomrule
  \endfoot
  
  \bottomrule
  \endlastfoot

  % SERAPHIC ORDER
  Inaea (Webs, Fate) & Persuade &
  \emph{Grants the hearth's warmth---shelter for the weary and mercy for the lost.} \\
  
  Oath of Flame \& Light (Dawn, Vows) & Command &
  \emph{Grants the fire of dawn, a vow that shields the faithful and sears the faithless.} \\
  
  Mykkiel (Zeal, Judgment) & Command &
  \emph{Grants the blazing verdict: covenant carried in fire and oath, binding law to blade.} \\
  
  Sacred Geometry (Order, Form) & Tinker &
  \emph{Grants the compass of perfection, reducing all shapes to their true measure.} \\
  
  Inquisitor Prime (Purity, Domination) & Insight &
  \emph{Grants the hunter's sight, piercing illusion and revealing supernatural corruption.} \\
  
  Gallows Bell (Justice, Reckoning) & Command &
  \emph{Grants the toll of ending, a voice that closes stories and calls all debts due.} \\
  
  Varnek Karn (Death, Arbitration) & Command &
  \emph{Grants the silence of the archive, where the dead obey and records speak.} \\
  
  Confessor Beneath the Bell (Burden, Exchange) & Command &
  \emph{Grants the weight of absolution, a burden shared between sinner and witness until the bell tolls.} \\
  
  Silent Choir (Mercy, Silence) & Insight &
  \emph{Grants the mercy of withheld truth---the stillness that spares and erases, leaving peace where confession would burn.} \\
  
  The Witness (Truth, Revelation) & Notice &
  \emph{Grants the unblinking gaze that unmasks deceit and remembers every oath.} \\
  
  Sealed Gate (Boundaries, Protection) & Tinker &
  \emph{Grants mastery of thresholds---doors that yield or bar at your command.} \\

  % EMPYREAL ORDER
  Ikasha (Shadow, Secrets) & Stealth &
  \emph{Grants the hush between footsteps and the raven's omen at every threshold.} \\
  
  Raéyn (Sea, Storms) & Skirmish &
  \emph{Grants the sailor's fortune---winds that shift and storms that answer to will.} \\
  
  Thrysos (Revels, Ecstasy) & Performance &
  \emph{Grants the satyr's crown---frenzy that drowns sorrow in wine and freedom in abandon.} \\
  
  Mab (Courts, Glamour) & Persuade &
  \emph{Grants the mask of favour, a voice that bends courtiers and kindles desire.} \\
  
  Livaea (Seduction, Social Binding) & Persuade &
  \emph{Grants the lure of longing, beauty sharpened into power over hearts and alliances.} \\
  
  Palinode (Performance, Rapture) & Performance &
  \emph{Grants the breath of inspiration, turning craft into rapture that binds its audience.} \\
  
  Aveh (Freedom, Erasure) & Stealth &
  \emph{Grants the storm's passage: vanishing paths and freedom unbound by law or memory, at the cost of leaving nothing behind.} \\
  
  Morag the Hag (Bargains, Transformation) & Deception &
  \emph{Grants the cunning of cruel deals, revealing hidden costs and twisting fate through negotiation.} \\
  
  Grimmir (Wild Speaker, Seasons) & Survival &
  \emph{Grants the wisdom of root and stone, speaking the tongues of leaf and fang---the rhythm of planting and harvest.} \\
  
  Maelstraeus (Infernal Bargainer) & Persuade &
  \emph{Grants the contract's weight, every deal sealed in fire and shadow.} \\
  
  The Traveler (Ways, Roads) & Notice &
  \emph{Grants the open way, a restless compass and roads where none were marked.} \\

  % CHTHONIC ORDER
  Isoka (Serpents, Renewal) & Skirmish &
  \emph{Grants the serpent's coil: strength in sudden strike and wisdom in renewal.} \\
  
  Malachai (Curses, Corruption) & Command &
  \emph{Grants the shadow in salvation---a gift that devours from within.} \\
  
  Carrion King (Decay, Renewal) & Survival &
  \emph{Grants the feast of decay, where what is dead becomes seed for what yet lives.} \\
  
  Nidhoggr (Deep Time, Memory) & Skirmish &
  \emph{Grants the weight of ages---the strength of stone and the hunger of roots.} \\
  
  Khemesh (Abyss, Pressure) & Skirmish &
  \emph{Grants the crushing silence of the deep, where strength drowns beneath weight.} \\
  
  Clockwork Monad (Iteration, Technology) & Tinker &
  \emph{Grants the certainty of repetition: a cycle that never falters, a gear that never slips.} \\
  
  Pale Shepherd (Thresholds, Guidance) & Stealth &
  \emph{Grants the shepherd's whisper, guiding the living through liminal spaces and comforting the dying.} \\
  
  Vorthak (Consumption, Hunger) & Stealth &
  \emph{Grants the endless hunger, devouring substance and meaning alike.} \\
  
  The Ninth (Void, Erasure) & Stealth &
  \emph{Grants dominion over absence, power found in silence and erasure.} \\
  
  Mor'iraath (Destruction, Unmaking) & Skirmish &
  \emph{Grants the entropic touch, seeing weakness and consuming value in ruin and flame.} \\
  
  Ninth Rim (Mouth of Silence) & Stealth &
  \emph{Grants the final erasure, where memory and self dissolve into the void.} \\

  % BEYOND THE TRIUNE
  Breath of the First Forge (Creativity, Ingenuity, Joy of Making) & Tinker &
  \emph{Grants the spark of creation: a fleeting inspiration that ignites the hand and warms the forge.} \\

  Kuva (Sky, Storms, Witness) & Notice &
  \emph{Grants the witness of the open dome: every promise made under sky is binding, and the weather answers those who watch.} \\
  
  Lucky Jack (Luck, Heist) & Deception &
  \emph{Grants the gambler's edge: a coin that never lands against you, a plan that always has one more twist.} \\
  
  Lunara (Moon, Mystery) & Notice &
  \emph{Grants the silver quiet, revealing seams and reflections that betray what is hidden.} \\
  
  Oya (Wind of the Sahel, Liberation) & Stealth &
  \emph{Grants the harmattan's breath: chains loosen, evidence scatters, and the wind remembers those who cannot be found.} \\
  
  Riven Path (Passion, Power) & Intimidation &
  \emph{Grants the breaking wave: chains loosen, wills shatter, and power flows to those who seize it.} \\
  
  Savage Heart (Primal Consumption) & Survival &
  \emph{Grants the red maw---the strength of bear and the hunger of the wolf. Predation without malice, survival without mercy.} \\
  
  Seal-Breaker (Forgotten Names, Erasure) & Deception &
  \emph{Grants the art of erasure: names fall from ledgers, seals crack without witness, and the Hollow collects the silence.} \\
  
  Unbroken Way (Balance, Control) & Insight &
  \emph{Grants the still point: clarity in chaos, calm in crisis, and the open hand that redirects force rather than meeting it.} \\
  
  Xhak'Thul (Storm, Shock, Awe) & Command &
  \emph{Grants the voice that splits mountains, the drum that shakes foundations with primal storm.} \\
\end{longtable}

\clearpage

\subsection{Thiasos Examples — Familiar Spirits by Patron}
\label{subsec:thiasos-table}

For Runekeepers, the Thiasos is a fragment of the Patron's attention given form. The following table offers typical manifestations. For Invokers, these serve as models for the kind of attention each Patron prefers—the shape of their presence in the world.

\begin{longtable}{@{}p{4.2cm}p{9.4cm}@{}}
  \toprule
  \textbf{Patron} & \textbf{Typical Thiasos} \\
  \midrule
  \endfirsthead
  
  \toprule
  \textbf{Patron} & \textbf{Typical Thiasos} \\
  \midrule
  \endhead
  
  \midrule
  \multicolumn{2}{r}{\emph{Continued on next page}} \\
  \bottomrule
  \endfoot
  
  \bottomrule
  \endlastfoot

  % SERAPHIC
  Inaea & A garden spider that weaves its web in the shape of your name; a small silk moth that leaves a trail of thread. \\
  Oath of Flame \& Light & A fire-coloured salamander that basks on hot coals; a tiny phoenix fledgling that glows faintly in darkness. \\
  Mykkiel & A white hound with brass eyes that can smell lies; a bronze-feathered hawk that circles above suspected heretics. \\
  Sacred Geometry & A brass beetle whose carapace is etched with golden-ratio spirals; a set of Kon'reh pieces that click advice. \\
  Inquisitor Prime & A white hound with brass eyes; a ferret that hunts rodents of corruption; a mechanical bird that sings hymns. \\
  Gallows Bell & A gray mouse that lives in courthouses; a brass-bell frog whose croak sounds like a distant chime. \\
  Varnek Karn & A lead seal that has witnessed a funeral; a knucklebone from an ancestor who died content. \\
  Confessor & A gray mouse that nests in church walls and listens to secrets; a bell-voiced cricket that chirps when a confession is true. \\
  Silent Choir & A gray mouse that nests in letter-boxes and never cheeps; a moth whose wings are dusted with forgetfulness. \\
  The Witness & A raven that collects shiny objects and recites forgotten names; a silverfish that eats lies from parchment. \\
  Sealed Gate & A bronze key that fits no lock yet trembles when a boundary is near; a brass bell that rings when a ward is about to break. \\

  % EMPYREAL
  Ikasha & A black cat with silver eyes; a gray mouse that nests in confessionals; a raven chick with one white feather. \\
  Raéyn & A storm-petrel that appears before gales; a gray seal pup that sings in harmonic tones. \\
  Thrysos & A satyr-kid with too-bright eyes; a fox that dances on its hind legs when the moon is full. \\
  Mab & A talking cricket that gives advice in riddles; a cat with mismatched eyes that appears at masquerades. \\
  Livaea & A white cat that curls around the necks of sleeping lovers; a red songbird that mimics a satisfied sigh. \\
  Palinode & A conductor's baton carved from a stage that was never completed; a silver bell that rings only for a perfect performance. \\
  Aveh & A storm-petrel that appears before gales; a gray fox that leaves no tracks; a shadow-hound pup whose fur shifts like cloud. \\
  Morag & A gray mouse that nests in counting-houses; a raven that steals shiny objects as bargaining chips. \\
  Grimmir & A pine marten with eyes like amber; a raven that knows the migration routes of every bird. \\
  Maelstraeus & A rat that counts coins with its paws; a ferret that steals only the most valuable trinkets. \\
  The Traveler & A gray hare that never tires; a sparrow that knows every shortcut; a tiny brass compass that hums on the true path. \\

  % CHTHONIC
  Isoka & A snake that sheds its skin weekly; a gecko that regrows its tail; a salamander that thrives in ashes. \\
  Malachai & A black rat with one red eye; a crow with a broken feather that points toward the nearest corpse. \\
  Carrion King & A carrion beetle that clicks its mandibles in a language of decomposition; a pale fungus that pulses when a death is near. \\
  Nidhoggr & A blind salamander that lives in caves; a fossilised beetle that occasionally twitches; a snail whose shell records years. \\
  Khemesh & A vial of abyssal water that never warms; a piece of coral that grows only in absolute darkness. \\
  Clockwork Monad & A brass-plated beetle that adjusts its carapace; a tiny clockwork lizard that repairs its own gears. \\
  Pale Shepherd & A gray mouse that nests in graveyards; a moth with dusty wings that lands on the hands of the dying. \\
  Vorthak & A glossy black beetle that devours rot; a pale grub that pupates only after consuming something precious. \\
  The Ninth & A sentient compass that points to contradictions; a brass abacus whose beads click in impossible sequences. \\
  Mor'iraath & A blackened salamander that sleeps in hot coals; a fire-scorpion whose sting leaves a cauterised wound. \\
  Ninth Rim & A fragment of polished obsidian that shows no reflection; a bell that tolls only in silence. \\

  % BEYOND
  Kuva & A storm petrel that nests on highest peaks; a wind-chaser beetle that clicks its shell when a gale is coming. \\
  Lucky Jack & A magpie that collects shiny objects; a mouse that nests in counting-houses and nibbles unpaid bills. \\
  Lunara & A silver-eyed owl that flies only at moonrise; a pale moth that emerges only under the full moon. \\
  Oya & A sand-coloured gecko that sheds its skin twice as fast; a small horned viper that leaves no tracks in loose sand. \\
  Riven Path & A cracked crystal that burns warm when anger rises; a shard of obsidian that hums with potential. \\
  Savage Heart & A wolf pup with amber eyes; a lynx kitten whose fur prickles at danger; a viper that coils around the wrist. \\
  Seal-Breaker & A ghost moth that appears only when a sealed door is opened; a clay scarab that crumbles when a false oath is sworn. \\
  Unbroken Way & A focus crystal that hums faintly when at peace; a translucent stone that grows cold when danger is near. \\
  Xhak'Thul & An electric blue salamander that crackles when touched; a tiny hawk whose feathers hum with static. \\
  Breath of the First Forge & A small, curious construct of living cogs and glowing embers---a shape-shifting presence that chatters with the sound of hammer on anvil and delights in showing its keeper novel ways to use familiar tools. \\
\end{longtable}

\clearpage

\subsection{Patron Rivalries — Cross-Resonance and Friction}
\label{subsec:rivalry-table}

When an Invoker carries Symbols from rival Patrons, Cross-Resonance generates friction. The table below details the nature of that friction and provides Story Beat prompts. \emph{Edge Loci} are environments where one side tends to start a step better in Position or gains an Effect nudge. \emph{Friction} are handy prompts for SB spends.

\renewcommand{\arraystretch}{1.15}
\setlength{\LTpre}{0pt}
\setlength{\LTpost}{0pt}

\begin{longtable}{@{}p{3.3cm}p{3.3cm}p{4.6cm}p{7.2cm}@{}}
  \toprule
  \textbf{Patron} & \textbf{Rival} & \textbf{Edge Loci} & \textbf{Friction \& Prompts (SB)} \\
  \midrule
  \endfirsthead

  \toprule
  \textbf{Patron} & \textbf{Rival} & \textbf{Edge Loci} & \textbf{Friction \& Prompts (SB)} \\
  \midrule
  \endhead

  \midrule
  \multicolumn{4}{r}{{Continued on next page}} \\
  \endfoot

  \bottomrule
  \endlastfoot

  % SERAPHIC RIVALRIES
  Aveh (Freedom, Erasure) & Oath of Flame \& Light (Dawn, Vows) &
  Crossroads, abandoned gates, wind-scoured battlefields. &
  SB: vows dissolve in the storm; liberty shatters oaths; the dawn sears absence into memory. \\

  Aveh & Varnek Karn (Death, Arbitration) &
  Forgotten cairns, erased ledgers, storm-blown libraries. &
  SB: records fade into silence; the storm devours testimony; archive cannot hold the exiled. \\

  Aveh & Sealed Gate (Boundaries, Closure) &
  Crumbling thresholds, broken wards, breached tombs. &
  SB: the Rider tramples seals; closure collapses to exile; freedom births chaos beyond boundary. \\

  Aveh & Carrion King (Decay, Renewal) &
  Open steppes, plague winds, unmarked graves. &
  SB: storm scatters the feast; renewal has no soil to root; liberty dissolves the cycle into absence. \\

  Aveh & Gallows Bell (Justice, Judgment) &
  Storm-dark tribunals, deserted execution grounds. &
  SB: judgment unheard in storm's roar; rope snaps in gale; freedom mocks justice, leaving silence. \\

  Breath of the First Forge & Clockwork Monad &
  Forges, workshops, design studios. &
  SB: perfection chokes inspiration; the spark stutters under endless refinement; the gear grinds the nascent idea. \\

  Breath of the First Forge & Oath of Flame \& Light &
  Altars, consecrated fires, vow-stones. &
  SB: the flame demands purity; the Breath craves experimentation; the vow sears the playful spark. \\

  Breath of the First Forge & Seal-Breaker &
  Forgotten ledgers, shattered seals, silent archives. &
  SB: the Breath builds upon fragments; the Seal-Breaker unbuilds what the Breath has just made; creation meets erasure. \\

  Breath of the First Forge & The Witness &
  Tribunal halls, truth-telling circles, places of record. &
  SB: the Witness demands accountability; the Breath rewards the attempt; truth can cage the joyful spark. \\

  Breath of the First Forge & Savage Heart &
  Hunting grounds, sacrificial altars, places of primal consumption. &
  SB: the Maw consumes the unfinished; the Breath cherishes the half-made; hunger clashes with the joy of making. \\

  Carrion King & Varnek Karn &
  Catacombs, bone-fields, ossuary halls. &
  SB: the feast of decay disputes record; skulls reorder themselves; scavengers mock the archive. \\

  Clockwork Monad & The Traveler &
  Caravans, milestones, repeating relays. &
  SB: schedule collapses; detour offers safety; routine blinds to ambush. \\

  Confessor & Oath of Flame \& Light &
  Burnt confessionals, sunrise absolutions, bells that toll at false dawn. &
  SB: vows cleanse through pain; guilt burns but leaves ash; the penitent's light blinds the absolver. \\

  Confessor & The Witness &
  Tribunal halls, bloodied ledgers, whispered testimonies. &
  SB: confession births revelation; truth festers in silence; guilt migrates from sinner to judge. \\

  Confessor & Silent Choir &
  Shuttered sanctuaries, wordless absolutions, drowned bells. &
  SB: mercy silences guilt before forgiveness is granted; silence absolves but leaves the soul hollow. \\

  Confessor & Sealed Gate &
  Locked confessionals, sealed reliquaries, hidden sins. &
  SB: penance denied; the gate holds sin too tightly; absolution festers within confinement. \\

  Gallows Bell & Mykkiel &
  Courts, tribunals, execution grounds. &
  SB: judgment pronounced but precedent denied; sentence demands haunting; writ echoes in the toll. \\

  Grimmir & Inquisitor Prime &
  Hedge-rows, forest shrines, contested boundaries between field and wood. &
  SB: the wild refuses the flame; order chokes root and vine; persecution kindles a green wrath. \\

  Grimmir & Carrion King &
  Rotting groves, compost-heaps, seasonal decay. &
  SB: decay feasts where the wild weeps; renewal becomes carrion; the season turns on itself. \\

  Ikasha & Oath of Flame \& Light &
  Twilight thresholds, oath-fires, vigil sites. &
  SB: light corners a secret; vow drags shadow forth; escape costs more than silence. \\

  Ikasha & The Witness &
  Silent cloisters, shaded rooms, hidden alcoves. &
  SB: hush worsens Position; shadows betray your step; deferred answers arrive with weight. \\

  Inaea & Livaea &
  Feasting halls, chambers of intimacy, hearth-feasts. &
  SB: desire strains hospitality; passion corrodes mercy; kindness is answered with demand. \\

  Inquisitor Prime & Morag the Hag &
  Crossroads, trial fires, cursed markets. &
  SB: bargain taints purification; soul scoured clean but scarred; purity test twists to hunger. \\

  Isoka & Sealed Gate &
  Doorways of shed skin, ritual molting chambers, thresholds of identity. &
  SB: the shed skin slips the gate; closure cannot hold what has already left; the serpent mocks the seal. \\

  Isoka & Oath of Mercy and Grace &
  Sickbeds where change is refused, hospices of last rites. &
  SB: mercy seeks to heal; Isoka sheds the wound whole; forgiveness becomes a second skin. \\

  Khemesh & Raéyn &
  Lightless holds, trenches, flooded caverns. &
  SB: abyss groans; storm maps fail; pressure crushes speech; tide demands oath. \\
  
  Kuva & Khemesh &
  Ship decks during storms, high peaks above deep chasms, places where sky meets sea. &
  SB: the dome witnesses, the abyss swallows; lightning illuminates the deep, but pressure drags the light down. \\
  
  Kuva & Sealed Gate &
  Unroofed thresholds, doorways open to the sky, bridges without walls. &
  SB: the gate bars passage; the sky demands witness; closure cannot seal what is already witnessed by the dome. \\
  
  Kuva & The Witness &
  Hilltop tribunals, open-air courts, storm-lit confessions. &
  SB: two witnesses compete: one records facts, the other remembers oaths; the sky judges where ink fails. \\

  Livaea & Maelstraeus &
  Masked balls, deal-feasts, salons. &
  SB: desire sweetens the bargain; pact sealed in touch not ink; temptation costs more than coin. \\

  Lucky Jack & Maelstraeus &
  Gambling dens, counting-houses, deal-tables. &
  SB: luck bends the contract; the Infernal Bargainer's fine print crumbles against a lucky roll; the coin spins until someone loses. \\

  Lucky Jack & Gallows Bell &
  Courtrooms, execution grounds, places of final judgment. &
  SB: justice demands a verdict; luck offers a reprieve; the Bell tolls but the coin is still in the air. \\

  Lucky Jack & The Witness &
  Interrogation rooms, truth-telling circles, places where secrets are exposed. &
  SB: luck hides the truth; the Witness demands it; a fortunate slip of the tongue or a misplaced piece of evidence alters the outcome. \\

  Lunara & The Witness &
  Mirrored chambers, silver pools, moonlit halls. &
  SB: reflection shows harsher truth; seams shine; witness misses what Lunara reveals. \\

  Mab & Palinode &
  Masques, salons, theaters. &
  SB: glamour demands encore; truth offends mask; favour must be sung aloud; court erupts at song's climax. \\

  Maelstraeus & Oath of Flame \& Light &
  Oath-stones, consecrated doors, witness trees. &
  SB: vow sears loophole shut; bargain price due early; sacred contract becomes trap. \\

  Malachai & Oath of Flame \& Light &
  Desecrated shrines, dawn-lit altars, sacred vows broken. &
  SB: a vow sanctifies but inflicts pain; the flame purges yet leaves despair. \\

  Malachai & The Witness &
  Confessionals, false miracles, dark reliquaries. &
  SB: a blessing unmasks a curse; truth corrodes faith; relics mock salvation; the witness must name corruption. \\

  Morag the Hag & Maelstraeus &
  Markets, bargain-halls, crossroads. &
  SB: deal's true cost unveiled; haste breeds ruin; balance exacts more than offered. \\

  Mor'iraath & Sacred Geometry &
  Shattered temples, collapsing bridges, ruined cities. &
  SB: pattern dissolves into static; perfect form breaks; compass needle spins; order devoured by entropy. \\

  Mykkiel & The Witness &
  Courtrooms, sealed archives, oath-bound chambers. &
  SB: judgment meets unblinking truth; the writ cannot hide; the covenant bleeds under witness. \\

  Nidhoggr & The Witness &
  Ruins, ancestor shrines, truth-seeing rites. &
  SB: revelation awakens the old; myth proven or destroyed; ancestor demands speech. \\

  Oath of Mercy and Grace & Inquisitor Prime &
  Hospice doorways, plague villages, sites of public mercy. &
  SB: mercy accused of weakness; forgiveness seen as treason; the Inquisitor demands payment where the Oath accepts tears. \\

  Oath of Mercy and Grace & Varnek Karn &
  Gravesides, funerary crypts, deathbeds. &
  SB: the dead owed nothing but the Weepers weep anyway; Karn's ledger rejects tears as legal tender; forgiveness becomes a debt to the archive. \\

  Oath of Mercy and Grace & Silent Choir &
  Shuttered hospices, wordless vigils, places where mercy is performed in silence. &
  SB: mercy without witness fades; silence answers the spiral; the Choir forgives by forgetting; the Oath forgives by remembering the debt. \\

  Oath of Flame \& Light & Varnek Karn &
  Memorial crypts, plague pits, testaments. &
  SB: bone ledger demands payment; consecration threatens archive; unfinished oath resurrects feud. \\
  
  Oya & Inquisitor Prime &
  Desert courts, open-air prisons, sites of public flogging. &
  SB: the wind scatters the inquisitor's writ; chains rust in the harmattan; persecution chokes on dust. \\
  
  Oya & Sealed Gate &
  Unlocked thresholds, breached quarantines, desert border posts. &
  SB: the gate holds but the wind erodes; closure cannot seal what the storm has already opened; Oya laughs at locks. \\
  
  Oya & The Witness &
  Burned courthouses, vanished evidence, sand-covered ledgers. &
  SB: the witness records, but Oya scatters; truth buried under red dust; revelation arrives too late, erased by wind. \\

  Pale Shepherd & Varnek Karn &
  Graveyards, liminal crossroads, shrines of passage. &
  SB: shepherd offers peace, Karn demands duty; soul cannot rest; death's duty vs death's silence. \\

  Palinode & Sacred Geometry &
  Concert halls, ritual theaters, metered dances. &
  SB: form clashes with feeling; perfection dulls art; audience craves chaos in pattern. \\

  Palinode & Thrysos &
  Festival grounds, taverns, moots. &
  SB: song feeds revel; revel swallows song; crowd erupts beyond control; inspiration burns to mania. \\

  Raéyn & Oath of Flame \& Light &
  Consecrated decks, sunrise rites at sea. &
  SB: dawn sears back abyss; vow drags crew into peril; storm promises aid with debt. \\

  Riven Path & Unbroken Way &
  Crossroads of choice, meditation chambers, arenas of contest. &
  SB: passion shatters stillness; control chokes the breaking wave; the closed fist and open hand meet---one yields, one takes, neither forgets. \\

  Riven Path & Oath of Flame \& Light &
  Desecrated shrines, battlefields, sites of broken oaths. &
  SB: vows are chains to be broken; the dawn reveals the wound; passion answers duty with fire. \\

  Riven Path & Sealed Gate &
  Prison walls, sealed vaults, thresholds that deny passage. &
  SB: the gate exists to be shattered; closure is an insult; freedom has no lock. \\

  Sacred Geometry & Ninth Rim &
  Perfect temples, empty circles, still points of order. &
  SB: the pattern confronts absence; form trembles at the void; nothing is measured and found wanting. \\

  Sacred Geometry & Silent Choir &
  Anechoic chambers, silent scriptoria, libraries of frozen light. &
  SB: silence perfects form; the pattern prays without words; mercy becomes structure, structure becomes mercy. \\

  Savage Heart & Oath of Flame \& Light &
  Blood-soaked altars, sacred hunting grounds, sites of sacrifice. &
  SB: the flame judges but the Maw consumes; vows of mercy break on fangs; dawn reveals the kill. \\

  Savage Heart & Grimmir &
  Wilderness edges, territory lines, places where predator and prey meet. &
  SB: the wild speaks of balance; the Red Maw answers with hunger; the season turns red. \\

  Sealed Gate & Ninth Rim &
  Failing wards, thinning thresholds, sealed vaults. &
  SB: door rattles; ward flickers; void slips past; seal admits what it bars. \\

  Sealed Gate & The Traveler &
  Customs houses, locks, liminal crossings. &
  SB: ledger wrong; key refuses; toll extracted; desire path collapses to law. \\
  
  Seal-Breaker & The Witness &
  Erased ledgers, burnt archives, silent testimonies. &
  SB: the witness records, the seal-breaker erases; truth and oblivion contest every memory; the Hollow collects what both leave behind. \\
  
  Seal-Breaker & Mykkiel &
  Voided contracts, broken seals, courts without record. &
  SB: the writ binds but the seal-breaker cracks; law without witness is void; obligation erased from the ledger. \\
  
  Seal-Breaker & Varnek Karn &
  Unmarked graves, forgotten cairns, archives where names fade. &
  SB: the dead owe nothing to the forgotten; Karn's ledger holds empty pages; the Seal-Breaker speaks and the record simply stops. \\

  Silent Choir & The Witness &
  Echoing cloisters, confessionals with sealed lips, mirrors that reflect no mouth. &
  SB: truth yearns to speak but mercy stills the tongue; revelation shatters peace; silence becomes its own confession. \\

  Silent Choir & Inaea &
  Shuttered hearths, mourning halls, letters never sent. &
  SB: mercy divides; the hearth forgives through warmth, the Choir through forgetting; love dies in gentleness unspoken. \\

  Silent Choir & Varnek Karn &
  Catacombs of still air, funerary archives where the names are erased. &
  SB: the Choir hums among the bones; records fall mute; memory yields to rest; death and mercy share the same hymn. \\

  Silent Choir & Sacred Geometry &
  Anechoic temples, fractal vaults that swallow sound. &
  SB: silence perfects form; entropy refines design; mercy becomes architecture; the pattern prays without words. \\

  Silent Choir & Aveh &
  Ruined cloisters, desert shrines scoured by wind. &
  SB: silence liberates; erasure redeems; mercy untethers the self until nothing remains but peace. \\

  Thrysos & Carrion King &
  Feast halls, bonfires, festival graves. &
  SB: revel spills to carrion; death becomes dance; feast breeds seed of decay. \\

  Thrysos & Livaea &
  Bacchic revels, decadent salons, wine-feasts. &
  SB: revel loosens restraint; passion binds harder; temptation drunk becomes vow. \\

  Ninth Rim & The Witness &
  Forgotten vaults, failing records, silent ruins. &
  SB: truth swallows itself; record fades to dust; memory refuses recall. \\

  The Traveler & Clockwork Monad &
  Pilgrim routes, caravans, timetables. &
  SB: routine bends to detour; detour grants swiftness but ambush; schedule cannot hold. \\

  Unbroken Way & Riven Path &
  Meditation chambers, arenas, thresholds of choice. &
  SB: stillness meets the breaking wave; control denies passion; the open hand refuses to close, and the closed fist refuses to open. \\

  Unbroken Way & Malachai &
  Desecrated shrines, sites of fallen meditation, places where balance shattered. &
  SB: calm resists the curse; corruption stains the still point; the Way absorbs and redirects, but the Chained Angel's hunger is patient. \\

  Unbroken Way & Savage Heart &
  Wilderness edges, hunting grounds, places where predator meets prey. &
  SB: the Way deflects; the Maw consumes; balance and hunger circle each other---neither yields, neither forgets. \\

  Varnek Karn & Mykkiel &
  Audit halls, courts martial, ledger-vaults. &
  SB: writ binds restless; sentence offends precedent; archive haunts judgment. \\

  Vorthak & Sealed Gate &
  Wards, vaults, guarded thresholds. &
  SB: hunger devours lock; seal becomes meal; barrier broken by appetite. \\

  Xhak'Thul & Oath of Flame \& Light &
  Storm-lit altars, dawn-thunder battles, sites of primal oaths sworn under sky. &
  SB: lightning sanctifies but burns the vow-breaker; storm drowns the dawn-cry; thunder answers fire with its own judgment. \\

  Xhak'Thul & Raéyn &
  Storm-tossed seas, coastal cliffs, places where sky and water meet. &
  SB: lightning splits the waves; storm-queen and thunder-speaker contest the wind; the crew prays to both, and both demand a price. \\  
\end{longtable}

\paragraph{Quick Rulings for Rivalries}
\label{subsec:rivalry-quick}

\begin{itemize}
  \item \textbf{Position Nudge:} In a home locus, start one step better; in a rival locus, one step worse.
  \item \textbf{Effect Shade:} Where a Patron dominates, consider an Effect bump; where opposed, consider Limited Effect unless paid for.
  \item \textbf{Symbol Interference:} Carrying both sides' Symbols increases narrative noise: first ritual each scene may mark +1 Obligation (GM's call).
  \item \textbf{SB Colour:} When spending SB in these matchups, prefer suits that fit: Hearts (social), Spades (harm/escalation), Clubs (material cost), Diamonds (numinous disturbance).
\end{itemize}

% ============================================================================
% APPENDIX: INVOKER MECHANICS
% ============================================================================

\clearpage
\appendix
\chapter{Invoker Mechanics Reference}
\label{app:invoker-mechanics}

This appendix collects the mechanical systems relevant to Invoker play: Obligation, Borrowed Grace, the Symbol system, and the Invoker's Rites.

\section{Obligation}
\label{app:obligation}

Obligation represents the narrative weight of a Patron's attention. It is not a punishment but a \emph{story engine}---the sense that a Patron has invested, noticed, and expects relevance.

\subsection{Obligation Capacity}

Your \textbf{Obligation Capacity} is `Spirit + Presence`. This is the total Obligation you can carry across \emph{all} Patrons before your ledger begins to creak.

\begin{itemize}
  \item \textbf{At or below Capacity:} You are balanced. Your Patrons are paid.
  \item \textbf{Exceeding Capacity:} For each segment over Capacity, mark 1 Fatigue.
  \item \textbf{Double Exceeding:} Your Patrons send a \emph{representative}.
\end{itemize}

\subsection{Clearing Obligation}

Obligation is cleared through \textbf{meaningful engagement} with the Patron.

\begin{itemize}
  \item \textbf{Off-screen (Act of Service):} Clear 1 segment by performing a resonant service aligned to the Patron's themes.
  \item \textbf{On-screen (Focused Devotion talent, 3--5 XP):} Once per scene, spend 1 Boon to clear 1 segment.
\end{itemize}

\section{Borrowed Grace}
\label{app:borrowed-grace}

Once per scene, when you visibly display a maintained Symbol, you may spend 1 Boon or 1 Fatigue as an action to gain:

\begin{itemize}
  \item \textbf{Combat:} +1 die to your next melee or ranged attack
  \item \textbf{Social:} +1 Effect on a social or stealth action
  \item \textbf{Thematic:} +1 die on a skill roll aligned with the Patron's domain
\end{itemize}

If the Symbol is Compromised, Borrowed Grace works at --1 die.

\section{The Symbol System}
\label{app:symbol-system}

\begin{itemize}
  \item \textbf{Acquisition:} Each Symbol costs 4 XP.
  \item \textbf{Attunement:} You may carry up to four Symbols without penalty.
  \item \textbf{Cross-Resonance:} Carrying Symbols from rival Patrons creates friction. The first invocation of a scene costs +1 Obligation.
  \item \textbf{Crack the Seal:} Resolve a Rite instantly; mark double Obligation; the Symbol becomes Compromised.
\end{itemize}

\section{Invoker's Rites}
\label{app:invoker-rites}

Invokers access Rites through their Symbols. The Difficulty Value for a Rite is:

\[
\text{DV} = \max\!\big(\text{Obligation Cost} - \text{Spirit}, \, \text{Tier of the rite (Low 1, Standard 2, High 3)}\big)
\]

\subsection{Universal Rites}

The following Rites are available to any Invoker who carries the appropriate Symbol.

\subsubsection*{Rite of Road-Sense (The Traveler, Low)}
\textit{Obligation: 4}
\begin{itemize}
  \item Gain +1 die to avoid ambushes or delays.
  \item Create a Path Memory timer to ignore one instance of difficult terrain.
\end{itemize}

\subsubsection*{Rite of the Unlit Candle (Ikasha, Low)}
\textit{Obligation: 6}
\begin{itemize}
  \item Gain +1 die to Stealth or Subterfuge.
  \item Declare one minor detail ``forgotten'' by any observer.
\end{itemize}

\subsubsection*{Rite of Hearth-Thread Knot (Inaea, Low)}
\textit{Obligation: 5}
\begin{itemize}
  \item Bind yourself and a willing target with a sympathetic thread.
  \item Both gain +1 die to Aid each other.
\end{itemize}

\subsubsection*{Rite of the Lingering Trace (The Witness, Low)}
\textit{Obligation: 4}
\begin{itemize}
  \item Gain +1 die to Investigation.
  \item Ask one yes/no question about a connection.
\end{itemize}

\subsubsection*{Rite of the Lucky Coin (Lucky Jack, Low)}
\textit{Obligation: 4}
\begin{itemize}
  \item Reroll one die that came up 1.
  \item If the reroll is also 1, the GM gains 2 SB.
\end{itemize}

\subsubsection*{Rite of Burden Transference (The Confessor, Low)}
\textit{Obligation: 5}
\begin{itemize}
  \item Transfer one Condition, Fatigue, or 1 Harm from another target to yourself.
\end{itemize}

\subsection{Invoker's Talents}

\paragraph*{Patron's Symbol (4 XP)}
Acquire a Symbol for a specific Patron. This is the foundation of Invoker practice.

\paragraph*{Focused Devotion (3--5 XP)}
Once per scene, spend 1 Boon to clear 1 segment of Obligation from a Patron (cannot clear segments accrued in the same scene).

\paragraph*{Crack Specialist (6 XP)}
Reduce the Obligation cost of Crack the Seal by 1 (minimum +1).

\paragraph*{Symbol Resonance (4 XP)}
When using a Symbol, gain +1 die to the ritual roll.

\section{Patron Pseudonyms — A Cultural Gazetteer}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=On Names and Masks,breakable]
The Patrons are not gods. They do not demand worship, nor do they care what name you speak when you invoke them. But mortals are creatures of meaning, and meaning demands a name. Across the Amaranthine, the same Patron wears a thousand faces. A Linn skald knows Ikasha as Aeki, the Trickster of the Fjords. A Vhasian knight prays to Saint Isabeau of the Unseen Door. An Ecktorian cultist whispers to the Spinner in the Shadows. They are all Ikasha. They are all wrong. They are all right.

This table collects the names the Patrons wear in different cultures. Use them to flavor your world, to mislead your players, or to remind yourself that the Patrons are not what mortals think they are.

\textit{``The name you speak is not the name I answer to. But I will answer anyway, because you called. That is the price of speaking names. That is the weight of being heard.''}
— The Gray Wanderer
\end{tcolorbox}

\subsection{The Seraphic Order — Law, Structure, Mercy}

\begin{longtable}{p{3cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}}
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endfirsthead
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endhead
\bottomrule
\endfoot
\textbf{Inaea} & Saint Aeliana the Weaver & The Spinner in the Shadows & The Thread-Mother & Vefari (The Weaver) \\
\textbf{Oath of Flame} & Saint Armand of the Dawn & The Unquenched Flame & The Light-Keeper & Brandr (The Brand) \\
\textbf{Mykkiel} & Saint Michel the Justiciar & The Gilded Scale & The Oath-Keeper & Vörðr (The Warden) \\
\textbf{The Witness} & Saint Veronique of the Mirror & The Unblinking Eye & The Truth-Teller & Sannr (The True One) \\
\textbf{The Sealed Gate} & Saint Germain the Threshold & The Unbroken Seal & The Door-Warden & Hliðr (The Gate) \\
\textbf{Sacred Geometry} & Saint Thierry the Measurer & The Perfect Form & The Pattern-Spinner & Talr (The Counter) \\
\textbf{Inquisitor Prime} & Saint Adalard the Pure & The Iron Hand & The Purge-Father & Brennr (The Burner) \\
\textbf{Gallow's Bell} & Saint Roland the Judge & The Tolling One & The Reckoner & Kalla (The Caller) \\
\textbf{Confessor Beneath the Bell} & Saint Remy the Unburdened & The Sin-Eater & The Burden-Bearer & Byrðr (The Burden) \\
\textbf{Silent Choir} & Saint Blanchette of the Silence & The Unspoken Vow & The Quiet One & Þögn (The Silence) \\
\textbf{Varnek Karn} & Saint Humbert the Archivist & The Ledger-Keeper & The Bone-Teller & Minni (The Memory) \\
\bottomrule
\end{longtable}

\subsection{The Empyreal Order — Motion, Desire, Performance}

\begin{longtable}{p{3cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}}
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endfirsthead
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endhead
\bottomrule
\endfoot
\textbf{Ikasha} & Saint Isabeau of the Unseen Door & The Shadow's Hush & The Shadow-Spinner & \textbf{Aeki (The Trickster)} \\
\textbf{The Traveler} & Saint Jacques the Pilgrim & The Open Road & The Wayfarer & Vegr (The Road) \\
\textbf{Palinode} & Saint Celeste the Encore & The Unfinished Song & The Crowd's Breath & Skaldr (The Song-Maker) \\
\textbf{Livaea} & Saint Marguerite the Siren & The Velvet Mask & The Heart-Taker & Lopt (The Charmer) \\
\textbf{Mab} & Saint Reynard the Fox & The Gilded Jest & The Glamour-Queen & Glæfr (The Glamer) \\
\textbf{Thrysos} & Saint Bacchus the Reveler & The Horned Circle & The Wild One & Dýrr (The Drunkard) \\
\textbf{Morag the Hag} & Saint Heloise the Bargainer & The Crossroads Crone & The Price-Mother & Kona (The Hag) \\
\textbf{Aveh} & Saint Aubert the Unbannered & The Storm-Rider & The Erased One & Vindr (The Wind) \\
\textbf{Grimmir} & Saint Aldric of the Wood & The Root-Father & The Antlered One & Skógr (The Forest) \\
\textbf{Maelstraeus} & Saint Lucien the Reckoner & The Infernal Bargainer & The Ledger-Devil & Skuld (The Debt) \\
\bottomrule
\end{longtable}

\subsection{The Chthonic Order — Dissolution, Renewal, Endings}

\begin{longtable}{p{3cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}}
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endfirsthead
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endhead
\bottomrule
\endfoot
\textbf{Isoka} & Saint Seraphine the Shedder & The Coiled Heart & The Skin-Shedder & Naðr (The Serpent) \\
\textbf{Malachai} & Saint Marguerite the Accursed & The Gilded Serpent & The Curse-Giver & Galdr (The Curse-Singer) \\
\textbf{Carrion King} & Saint Lazare the Composter & The Rot-Thread & The Bone-Father & Rotr (The Rot) \\
\textbf{Nidhoggr} & Saint Ambroise the Deep & The World-Worm & The Dreaming Stone & Ormr (The Worm) \\
\textbf{Khemesh} & Saint Damien of the Deep & The Abyssal Maw & The Drowning One & Hrymr (The Drowning) \\
\textbf{Clockwork Monad} & Saint Julien the Iterative & The Optimization Engine & The Ticking Heart & Véttr (The Mechanism) \\
\textbf{The Pale Shepherd} & Saint Bernadette of the Crossing & The Gray Flock & The Threshold-Guide & Leiðir (The Guide) \\
\textbf{Vorthak} & Saint Thibault the Hungry & The Ravenous Maw & The Feast-Bringer & Hungr (The Hunger) \\
\textbf{The Ninth} & Saint Urbain the Erased & The Unfinished Index & The Void-Bringer & Tómr (The Void) \\
\textbf{Mor'iraath} & Saint Aethelred the Unmaker & The Ashen Hand & The Fire-Bringer & Brenna (The Burner) \\
\bottomrule
\end{longtable}

\subsection{Beyond the Triune — The Unclassifiable}

\begin{longtable}{p{3cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}|p{2.8cm}}
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endfirsthead
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Ecktorian (Cult/Saint)} & \textbf{Theona (Folk)} & \textbf{Linn (Skaldic)} \\
\midrule
\endhead
\bottomrule
\endfoot
\textbf{Xhak'Thul} & Saint Thibault the Thunderer & The Storm-Caller & The Sky-Father & Brimir (The Storm-Father) \\
\textbf{Kuva} & Saint Aelric of the Dome & The Sky-Witness & The Cloud-Mother & Himinn (The Sky) \\
\textbf{Lucky Jack} & Saint Renard the Lucky & The Unspent Coin & The Magpie-King & Hamingja (The Luck) \\
\textbf{Lunara} & Saint Luna the Quiet & The Silver Mirror & The Moon-Queen & Máninn (The Moon) \\
\textbf{Oya} & Saint Imani the Scatterer & The Red Dust & The Wind-Mother & Vindr (The Wind) \\
\textbf{The Riven Path} & — (Heretical) & The Unbound Circle & The Chain-Breaker & Bauti (The Striker) \\
\textbf{The Savage Heart} & Saint Morcant the Red & The Painted Maw & The Red Maw & Blóðr (The Blood) \\
\textbf{Seal-Breaker} & Saint Isembart the Unsealed & The Broken Seal & The Name-Flayer & Óþekktur (The Unknown) \\
\textbf{Unbroken Way} & Saint Benedict of the Balance & The Still Point & The Walker & Jafnaðr (The Balancer) \\
\bottomrule
\end{longtable}

\subsection{Cultural Masks — The Patrons by Region}

\begin{longtable}{p{3cm}|p{2.5cm}|p{3cm}|p{2.5cm}|p{2.5cm}}
\toprule
\textbf{Patron} & \textbf{Ykrul (Clan)} & \textbf{Vilikari (Hybrid)} & \textbf{Aeler (Breath)} & \textbf{Mistlands (Bell)} \\
\midrule
\endfirsthead
\toprule
\textbf{Patron} & \textbf{Ykrul (Clan)} & \textbf{Vilikari (Hybrid)} & \textbf{Aeler (Breath)} & \textbf{Mistlands (Bell)} \\
\midrule
\endhead
\bottomrule
\endfoot
\textbf{Ikasha} & Shadow-Kin & The Gray Guide & The Deep Listener & The Bell-Shrouded \\
\textbf{Mykkiel} & Oath-Keeper & The Seal-Bearer & The Stone-Counter & The Bell-Judge \\
\textbf{The Witness} & Truth-Teller & The Record-Keeper & The Memory-Keeper & The Fog-Scribe \\
\textbf{Inaea} & Weaver-Mother & The Hearth-Spinner & The Root-Weaver & The Kettle-Spinner \\
\textbf{Isoka} & Shed-One & The Skin-Shedder & The Serpent-Coil & The Scale-Mother \\
\textbf{Malachai} & Curse-Giver & The Silver-Tongued & The Chained One & The Salt-Fang \\
\textbf{The Traveler} & Way-Walker & The Road-Warden & The Ways-Stone & The Lantern-Bearer \\
\textbf{Carrion King} & Bone-Father & The Field-Father & The Deep Mold & The Fog-Mother's Consort \\
\textbf{Xhak'Thul} & Thunder-Hoof & The Thunder-Voice & The Storm-Heart & The Storm-Bell \\
\textbf{Khemesh} & Deep-Hunger & The Deep-Strider & The Pressure & The Drowned Chorus \\
\textbf{Palinode} & Song-Maker & The Story-Singer & The Echo-Singer & The Bellsong \\
\bottomrule
\end{longtable}

\subsection{The God of Trickery — A Special Note}

The Linn know the trickster as \textbf{Aeki}, a figure of mischief, shape-shifting, and the space between oaths. Aeki is not evil—they are simply unpredictable. They laugh when promises are broken and smile when doors open the wrong way. Aeki is the god of the missed step, the forgotten name, the coin that lands on its edge.

\textbf{Identity:} Aeki is a pseudonym for \textbf{Ikasha}, She Who Sleeps. But the Linn do not know this. They call her Aeki, and they tell stories of her cunning. She does not correct them. She is patient.

Other cultures know the trickster by different names:

\begin{center}
\begin{tabular}{p{3cm}|p{4cm}|p{4cm}}
\toprule
\textbf{Culture} & \textbf{Name} & \textbf{Patron} \\
\midrule
Linn & Aeki & Ikasha \\
Vhasia & Saint Isabeau of the Unseen Door & Ikasha \\
Ecktoria & The Shadow's Hush & Ikasha \\
Theona & The Shadow-Spinner & Ikasha \\
Ykrul & Shadow-Kin & Ikasha \\
Vilikari & The Gray Guide & Ikasha \\
Aeler & The Deep Listener & Ikasha \\
Mistlands & The Bell-Shrouded & Ikasha \\
\bottomrule
\end{tabular}
\end{center}

\subsection{The Vilikari and Crom}

Among the Vilikari, the name \textbf{Crom} (or \textbf{Krom}, depending on the dialect) is spoken with a mixture of reverence and dread. Crom is not a god of worship. He is a god of \textit{survival}. He gives you strength at birth and expects you to use it. He does not answer prayers. He does not reward the faithful. He \textit{watches}, and he \textit{judges}.

\begin{quote}
\textit{``Crom laughs at your four winds. He gives you strength and expects you to use it. He does not care who you are or what you believe. He cares only that you are strong enough to stand when the storm comes.''}
— Vilikari Proverb
\end{quote}

\textbf{Identity:} Crom is a pseudonym for \textbf{The Savage Heart (Red Maw)}. The Vilikari do not know this. They think Crom is a god of the old steppe, a remnant of the Ykrul past. They are wrong. The Savage Heart does not correct them. The Savage Heart is patient.

But some scholars whisper that Crom is not The Savage Heart at all. They argue that Crom is \textbf{Xhak'Thul}, the Thunder-Speaker, shaped by Vilikari pragmatism into a god of raw, indifferent strength. The debate continues. The Patrons do not weigh in.

\begin{center}
\begin{tabular}{p{3cm}|p{4cm}|p{4cm}}
\toprule
\textbf{Culture} & \textbf{Name} & \textbf{Patron (Debated)} \\
\midrule
Vilikari (Traditional) & Crom/Krom & The Savage Heart \\
Vilikari (Reformist) & Krom & Xhak'Thul \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Using Pseudonyms in Play}

\begin{itemize}
  \item \textbf{Disguise the Patron.} A player may not know that Aeki is Ikasha. They may swear by Aeki and think they are safe. The GM knows the truth.
  \item \textbf{Cultural Flavor.} A Vhasian knight prays to Saint Michel. A Linn skald calls on Vörðr. They are both invoking Mykkiel. The names change; the weight does not.
  \item \textbf{Misinformation.} A cult may worship a false aspect of a Patron, using a name the Patron does not recognize. The Patron may answer anyway, out of curiosity or amusement.
  \item \textbf{Syncretism.} In mixed cultures, the same Patron may have multiple names. A Vilikari traveler may call on Crom in the morning and the Thunder-Voice at night. Both answer.
  \item \textbf{The Hollow's Interest.} The Hollow does not have a name. It does not need one. But mortals have tried. The names are always wrong. The Hollow does not correct them. The Hollow is patient.
\end{itemize}

\begin{tcolorbox}[colback=green!5!white,colframe=green!70!black,title=The Gray Wanderer's Advice,breakable]
\textit{``Call them what you will. They will answer. The name is not the thing. The thing is the weight. And the weight is always the same.''}
\end{tcolorbox}

\subsection{Name Generation Notes}

\textbf{Vhasian Saints:} Norman-French influenced. Elegant, courtly, often martial. Examples: Michel, Isabeau, Armand, Veronique, Germain, Thierry, Adalard, Roland, Remy, Humbert, Jacques, Celeste, Marguerite, Reynard, Heloise, Aubert, Aldric, Lucien, Seraphine, Lazare, Ambroise, Damien, Julien, Bernadette, Thibault, Urbain, Aethelred, Aelric, Renard, Luna, Imani, Morcant, Isembart, Benedict.

\textbf{Ecktorian Cults:} Mystery cults with evocative, ominous names. Examples: The Spinner in the Shadows, The Unquenched Flame, The Gilded Scale, The Unblinking Eye, The Unbroken Seal, The Perfect Form, The Iron Hand, The Tolling One, The Sin-Eater, The Unspoken Vow, The Shadow's Hush, The Open Road, The Unfinished Song, The Velvet Mask, The Gilded Jest, The Horned Circle, The Crossroads Crone, The Storm-Rider, The Root-Father, The Infernal Bargainer, The Coiled Heart, The Gilded Serpent, The Rot-Thread, The World-Worm, The Abyssal Maw, The Optimization Engine, The Gray Flock, The Ravenous Maw, The Unfinished Index, The Ashen Hand, The Storm-Caller, The Sky-Witness, The Unspent Coin, The Silver Mirror, The Red Dust, The Unbound Circle, The Painted Maw, The Broken Seal, The Still Point.

\textbf{Theona Folk:} Irish-Celtic and faerie-tale inspired. Examples: The Thread-Mother, The Light-Keeper, The Oath-Keeper, The Truth-Teller, The Door-Warden, The Pattern-Spinner, The Purge-Father, The Reckoner, The Burden-Bearer, The Quiet One, The Bone-Teller, The Shadow-Spinner, The Wayfarer, The Crowd's Breath, The Heart-Taker, The Glamour-Queen, The Wild One, The Price-Mother, The Erased One, The Antlered One, The Ledger-Devil, The Skin-Shedder, The Curse-Giver, The Bone-Father, The Dreaming Stone, The Drowning One, The Ticking Heart, The Threshold-Guide, The Feast-Bringer, The Void-Bringer, The Fire-Bringer, The Sky-Father, The Cloud-Mother, The Magpie-King, The Moon-Queen, The Wind-Mother, The Chain-Breaker, The Red Maw, The Name-Flayer, The Walker.

\textbf{Linn Skaldic:} Old Norse inspired. Short, guttural, resonant. Examples: Aeki, Vefari, Brandr, Vörðr, Sannr, Hliðr, Talr, Brennr, Kalla, Byrðr, Þögn, Minni, Vegr, Skaldr, Lopt, Glæfr, Dýrr, Kona, Vindr, Skógr, Skuld, Naðr, Galdr, Rotr, Ormr, Hrymr, Véttr, Leiðir, Hungr, Tómr, Brimir, Himinn, Hamingja, Máninn, Bauti, Blóðr, Óþekktur, Jafnaðr.

\textbf{Ykrul Clan:} Steppe-nomad titles. Descriptive, pragmatic. Examples: Shadow-Kin, Oath-Keeper, Truth-Teller, Weaver-Mother, Shed-One, Curse-Giver, Way-Walker, Bone-Father, Thunder-Hoof, Deep-Hunger, Song-Maker.

\textbf{Vilikari Hybrid:} Blends Utaran and steppe influences. Examples: The Gray Guide, The Seal-Bearer, The Record-Keeper, The Hearth-Spinner, The Skin-Shedder, The Silver-Tongued, The Road-Warden, The Field-Father, The Thunder-Voice, The Deep-Strider, The Story-Singer, Crom/Krom.

\textbf{Aeler Breath:} Dwarven, practical, stone-oriented. Examples: The Deep Listener, The Stone-Counter, The Memory-Keeper, The Root-Weaver, The Serpent-Coil, The Chained One, The Ways-Stone, The Deep Mold, The Storm-Heart, The Pressure, The Echo-Singer.

\textbf{Mistlands Bell:} Folk-horror, bell-themed. Examples: The Bell-Shrouded, The Bell-Judge, The Fog-Scribe, The Kettle-Spinner, The Scale-Mother, The Salt-Fang, The Lantern-Bearer, The Fog-Mother's Consort, The Storm-Bell, The Drowned Chorus, The Bellsong.

\subsection{Quick Reference: Patron to Pseudonym}

\begin{longtable}{p{3cm}|p{2.5cm}|p{2.5cm}|p{2.5cm}|p{2.5cm}}
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Linn (Skaldic)} & \textbf{Vilikari (Hybrid)} & \textbf{Theona (Folk)} \\
\midrule
\endfirsthead
\toprule
\textbf{Patron} & \textbf{Vhasian (Saint)} & \textbf{Linn (Skaldic)} & \textbf{Vilikari (Hybrid)} & \textbf{Theona (Folk)} \\
\midrule
\endhead
\bottomrule
\endfoot
Ikasha & Saint Isabeau & Aeki & The Gray Guide & The Shadow-Spinner \\
Mykkiel & Saint Michel & Vörðr & The Seal-Bearer & The Oath-Keeper \\
The Witness & Saint Veronique & Sannr & The Record-Keeper & The Truth-Teller \\
Inaea & Saint Aeliana & Vefari & The Hearth-Spinner & The Thread-Mother \\
Isoka & Saint Seraphine & Naðr & The Skin-Shedder & The Skin-Shedder \\
Malachai & Saint Marguerite & Galdr & The Silver-Tongued & The Curse-Giver \\
The Traveler & Saint Jacques & Vegr & The Road-Warden & The Wayfarer \\
Carrion King & Saint Lazare & Rotr & The Field-Father & The Bone-Father \\
Xhak'Thul & Saint Thibault & Brimir & The Thunder-Voice & The Sky-Father \\
Khemesh & Saint Damien & Hrymr & The Deep-Strider & The Drowning One \\
Palinode & Saint Celeste & Skaldr & The Story-Singer & The Crowd's Breath \\
Savage Heart & Saint Morcant & Blóðr & Crom/Krom & The Red Maw \\
\bottomrule
\end{longtable}

\subsection{Final Note: The Patrons Do Not Care}

The Patrons do not care what name you speak. They do not care if you know their true name. They do not care if you think you are invoking a different power. They answer because you called. The name is a gesture. The weight is the thing.

\textit{``Speak my name. Speak the wrong name. Speak silence. It does not matter. The weight is the thing. The weight is always the thing.''}
— The Gray Wanderer