% --- Patron: Khemesh, the Abyssal Maw (Depths, Inexorability, Eldritch Terror) ---
\subsection{Khemesh, the Abyssal Maw (Depths, Inexorability, Eldritch Terror)}

\textit{Lore.} Khemesh is not merely a lord of the depths but the hunger beneath them, a pressure older than seas. Those who bargain with him are marked by the abyss — seen in the way shadows cling, in the whispers heard when no voice speaks, in the certainty that all things will sink. He is the classic warlock’s patron: distant, vast, and utterly indifferent to mortal morality. To serve Khemesh is to accept that the universe is cold, that the deep does not hate — it simply consumes.

\begin{quote}
In the trench without light, the Maw waits. Even silence drowns.
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Depths and Pressure:} crushing weight, darkness, the alien geometry of the abyss
  \item \textbf{Inexorability:} slow, unstoppable processes; the certainty of sinking
  \item \textbf{Eldritch Terror:} madness, whispers, perception beyond mortal limits
  \item \textbf{Consumption:} hunger that devours matter, memory, and hope
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Khemesh does not keep a living familiar. Their \textbf{Thiasos} is a \textit{preserved remnant of the deep}: a vial of abyssal water that never warms, a piece of coral that grows only in absolute darkness, or a polished tooth from a creature that has never seen the sun. This object pulses faintly when pressure changes. It requires no food, but it must be kept submerged in salt water when not in use. If left dry for a week, it becomes \textit{Compromised} (it emits a distressing keening at random, imposing −1 die on Stealth rolls for its bearer).

The \textbf{Codex} of Khemesh is never a book of paper. It may be a slab of slate etched with runes that weep condensation, a chain of iron links each inscribed with a single syllable, or the inside of a deep‑sea diving helmet. Because the deep resists being fixed in writing, the Codex requires \textit{upkeep}: the Runekeeper must immerse it in cold salt water for an hour each week, or the Rites stored there become illegible (treat as Compromised until restored).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Khemesh may touch a held item (weapon, shield, or tool) and whisper a promise of sinking. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Intimidation} or \textit{Endurance} rolls, and its first use each scene \textit{ignores one point of Armor} (the deep’s pressure finds the cracks). After the scene, the user marks \textbf{+1 Obligation} as the cold of the abyss lingers in their bones.

\subsubsection*{Rites of Khemesh}

\paragraph{Rite of the Crushing Whisper (Low, 5 XP)} \hfill \\
\emph{Instant/Scene; Near/Zone; Yes.} \\
\textbf{Tags:} \texttt{[FEAR]}, \texttt{[DEBUFF]}, \texttt{[AREA]}, \texttt{[SILENCE]} \\
\textbf{Materials:} A broken shell filled with ink‑dark water. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Whisper of the Trench:} Target hears impossible echoes and suffers \textbf{--1 die} on their next action.
  \item \textbf{Crushing Silence:} Establish an oppressive silence in a zone; enemies in the zone suffer \textbf{--1 die} to coordination or morale‑driven actions.
\end{itemize}
\textbf{Push It:} (Whisper) Echoes coil in your own skull — take \textbf{Fatigue 1}, but the target also loses their next minor action. (Silence) A single enemy’s voice is stolen entirely for the scene. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Abyssal Grasp (Standard, 7 XP)} \hfill \\
\emph{Instant; Near; No.} \\
\textbf{Tags:} \texttt{[BIND]}, \texttt{[PRESSURE]}, \texttt{[DEBUFF]} \\
\textbf{Materials:} A length of rusted chain submerged in water. \\
\textbf{Effect:} Target is pinned by invisible crushing force: treat as \texttt{[ENTANGLE]} with \textbf{Great Effect} if underwater or confined. \\
\textbf{Push It:} Inflict \textbf{Fatigue 1} on the target in addition to the restraint. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Alien Gaze (Standard, 9 XP)} \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[BURDEN]}, \texttt{[FEAR]} \\
\textbf{Materials:} None (requires standing in absolute darkness or near deep water). \\
\textbf{Effect:} You perceive the world as Khemesh does — fractured, alien, crushing. Gain \textbf{+2 dice} to \textit{Notice} and \textit{Arcana}, and may ask one “true nature” question about a foe or structure. \\
\textbf{Cost:} When the scene ends, you suffer \textbf{Exposure +1} as your perception warps. \\
\textbf{Push It:} Extend the vision to one ally, but both take \textbf{Fatigue 1}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Sundering Deep (High, 14 XP)} \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[PRESSURE]}, \texttt{[AREA]}, \texttt{[FATE]}, \texttt{[DEATH]} \\
\textbf{Materials:} A sealed vessel of abyssal water, broken open. \\
\textbf{Effect:} Reality in the zone folds inward like the crushing deep:
\begin{itemize}
  \item Enemies act at \textbf{Desperate Position} by default.
  \item Each beat, the GM may force \textbf{1 SB} (Spades/Clubs favoured) as the deep pressures reality.
  \item Structures, vessels, or wards fracture as if under immense weight.
\end{itemize}
\textbf{Push It:} For one beat, declare a single enemy “crushed” (severe harm/effect). You immediately suffer \textbf{Fatigue 2} and \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} 1 action; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 8 segments.

\subsubsection*{Cantors \& Cults of Khemesh}

\textbf{The Cantor's Song.} A Cantor of Khemesh sings in \textit{pressures} — low, resonant drones that feel like the weight of water on the chest. Their voice is used to invoke the crushing depths, to rupture a ship's hull with a single sustained note, or to fill a room with the sensation of drowning. Khemesh's Cantors are rare, feared, and often found in coastal cults where they sing the tides into madness.

\textbf{The Cult of the Drowned Chorus.} Khemesh's cults are the stuff of eldritch horror — sunken temples beneath storm‑tossed seas, congregations of shore‑dwellers with wet eyes and too‑pale skin, and the occasional deep one or kraken‑spawn acolyte. A ritual of the Drowned Chorus involves each member cutting a small sigil into their palm (a spiral within a triangle) and then drinking a cup of brine and salt water while whispering a secret hope into the cup — hope that the depths will grant, or consume. The water \textit{whispers back} in a voice that sounds like their own, but wrong. Their creed is whispered in wet, gurgling tones: \textit{"All things sink. The only question is what you carry down with you."} Outsiders are bait, or converts. Sometimes both.

\subsubsection*{Witchcraft of Khemesh (Lay / Pressure)}

A witch who walks with Khemesh is a \textit{deep‑salvager} or \textit{pressure‑scribe} — one who works with the spoils of the abyss. Their craft is the art of enduring pressure: sealing a letter with wax drawn from deep‑sea vents, inscribing a ward that holds back the tide, or writing a contract in ink made from the blood of a drowned sailor. Their tools are not hammers; they use vials, seals, and the patient weight of water. In Linn, a shipwright might season a keel with a prayer, but a hedge‑witch of Khemesh might instead seal a curse into the hull with a drop of abyssal water, so that the ship will never sink — but its crew will dream of drowning every night.

\textbf{The Witch’s Tool: The Vessel of the Unspoken Depth.} A small, sealed iron vial, cold enough to frost the skin. Inside is a single drop of water from a trench where no light reaches. Once per session, the witch may uncap the vial and speak a name. For the next hour, that person (or object) is “marked by the deep”: they suffer \textbf{--1 die} on all social rolls (their presence feels oppressive), and any attempt to track them gains \textbf{+1 die} as if following a trail of cold. The vial can also be used to “seal” a threshold: pouring a single drop onto a doorframe causes the door to groan with pressure; forcing it open requires a Body test (DV 4) and inflicts 1 Fatigue on the opener.

\textbf{A Hedge Gift: Carried Depth (2 XP).} Once per scene, the witch may touch a container (a bottle, a chest, a room) and fill it with the “pressure of the deep.” For the next hour, any living creature inside that container must test Endurance (DV 3) or suffer 1 Fatigue from the crushing weight. The container may also be used to preserve food or seal a wound (the pressure stops bleeding), but the flesh grows pale and cold.

\subsubsection*{Khemesh's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Abyssal Resilience:} +1 die to resist fear and pressure‑based effects. & \textbf{Claustrophobic Comfort:} Suffer --1 die in open, well‑lit spaces or above ground. \\
2 & \textbf{Crushing Insight:} Once/scene, treat a failed Investigation or Arcana roll as a success, but mark 1 SB (Clubs). & \textbf{Weight of Knowledge:} When you learn a secret someone wanted hidden, test Resolve (DV 3) or gain 1 Fatigue as the abyss presses on you. \\
3 & \textbf{Silent Hunter:} +2 dice to Stealth in dark or confined spaces. & \textbf{Voice of the Deep:} When speaking normally, your voice sounds distant and hollow, causing --1 die to social rolls requiring warmth or clarity. \\
4 & \textbf{Pressure Adaptation:} Immune to underwater combat penalties; gain +1 die to resist drowning. & \textbf{Crushing Presence:} Allies within Near range suffer --1 die to morale‑based rolls due to your oppressive aura. \\
5 & \textbf{Abyssal Empathy:} Once/session, sense the deepest fear of any creature within Near range (GM tells you one thing they are afraid of). & \textbf{Fractured Perception:} --1 die to rolls requiring normal vision; the world appears warped and alien. You cannot recover Fatigue from rest while at sea. \\
6+ & \textbf{Maw-Touched:} Once/session, extinguish all non‑magical light sources in a zone as a free action. Your presence causes light to dim. & \textbf{Hunger of the Maw:} Mark +2 Obligation when using this power. You are vulnerable to fire (take +1 Harm from fire‑based attacks). Each dawn, you must consume something (food, memory, hope) or mark 1 Fatigue. \\
\end{tblr}

\subsection*{Playstyle Notes}
Khemesh is the classic warlock’s patron: distant, vast, and utterly indifferent. His followers gain power over pressure, darkness, and the crushing inevitability of the deep, but pay with their warmth, their humanity, and eventually their sanity. The cults of Khemesh are eldritch and alien, and even the most benevolent follower of the Maw is marked by a cold that never leaves. Ideal for players who enjoy cosmic horror, the price of forbidden knowledge, and characters who have looked into the abyss — and let it look back.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Eldritch Horror:} powers that are alien, vast, and morally indifferent
  \item \textbf{Crushing Pressure:} mechanical weight, slowness, inevitability
  \item \textbf{Deep as Warlock Patron:} bargains with the abyss, forbidden knowledge, madness
  \item \textbf{Consumption:} hunger that devours matter, memory, and hope
  \item \textbf{Isolation:} the follower grows cold, distant, and unsettling to others
\end{itemize}