% −−− Patron: Kuva (Sky, Storms, Witness of Oaths) −−−
\subsection{Kuva — The Sky That Takes Many Names}

\textit{Lore.} Kuva is the dome above all roads — the endless blue, the monsoon wrath, the night’s silent gaze. To the Kuvani of the eastern steppe, Kuva is a stern witness who hears every promise made under open heaven. To the Dhaharan of the fertile lowlands, Kuva is a storm‑mother, destroyer and renewer, whose rains demand sacrifice and give life. To the Tulkani wanderers, Kuva is a shape‑changing presence — neither kind nor cruel, only watchful, and a slighted sky brings ill fortune. To the Aelinnel, Kuva is the first counter, the one who taught the counting song. To the Aelaerem, Kuva is the Hollow’s boundary — the sky is where the Hollow’s gaze stops.

Kuva has no single shrine because Kuva is the roof of every shrine. When a Kuvani pours milk upward, when a Dhaharan offers blood at a monsoon bell, when a Tulkani ties a knot and throws ash, Kuva watches. It does not correct. It remembers.

\begin{quote}
``Kuva does not answer in words. It answers in weather. A clear day means ‘go.’ A storm means ‘wait.’ Lightning means ‘run.’ Learn to read the sky, and you will never need a priest.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  ℹtem \textbf{Witness of Oaths:} promises made under open sky are binding
  ℹtem \textbf{Storm & Weather:} rain, lightning, wind, and the monsoon’s rhythm
  ℹtem \textbf{Sky as Ledger:} the dome records every deed, every lie, every breath
  ℹtem \textbf{Threshold of the Hollow:} where the sky’s gaze ends, the Hollow begins
\end{itemize}

⫕section*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Kuva binds a creature of the open air. Their \textbf{Thiasos} is a small beast or mythological being that embodies wind and witness: a storm petrel that nests only on the highest peaks, a wind‑chaser beetle that clicks its shell when a gale is coming, a grey‑feathered kestrel that never lands on a roof, a small silver moth that appears only during lightning storms, or a field mouse that sits perfectly still when the sky is about to break. This creature requires no food, but it must be shown the open sky each day — a glimpse through a window is not enough. If confined for a week, it becomes \textit{Compromised} (its feathers or scales grow dull, and its bearer suffers –1 die on all navigation and weather‑sense rolls until it is released).

The \textbf{Codex} of Kuva is never a book kept under a roof. It may be a bundle of wind‑knotted cords that must be untied and retied each dawn, a slate tablet that must be read in the rain, a scroll of vellum that shrivels if not exposed to the elements, or a leather pouch of stones collected from nine different hilltops. Because the sky resists being fixed in ink, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week under open sky, reciting each Rite aloud to the wind, or the words fade (treat as Compromised until re‑consecrated in a storm).

⫕section*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Kuva may touch a held item (weapon, tool, or garment) and whisper a single word of the sky. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Navigation} or \textit{Survival} rolls made outdoors, and the first time each scene the bearer would be surprised by a weather‑related hazard (lightning, flash flood, sudden fog), they may reroll the resistance check (the sky warns its own). After the scene, the user marks \textbf{+1 Obligation} as a thread of sky‑debt tethers them to the open dome.

⫕section*{Rites of Kuva}

\paragraph{Rite of the Sky’s Witness (Low, 4 XP)} ℎfill \\
\emph{Scene; Near; No} \\
\textbf{Tags:} \texttt{[OATH]}, \texttt{[TRUTH]}, \texttt{[FEAR]} \\
\textbf{Materials:} A pinch of ash thrown upward, or a single breath spoken into a knot. \\
\textbf{Effect:} For one scene, when you speak a promise outdoors, any listener who hears it must test \textit{Resolve} (DV 3) to lie about that promise. If you break your own promise, the GM gains 2 Story Beats to spend on a sky‑related complication (lightning, sudden fog, a hawk stealing your food). \\
\textbf{Push It:} The oath is witnessed by a passing storm — the liar suffers –2 dice to all rolls until they make amends under open sky. Mark \textbf{1 SB (Diamonds)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of Kuva’s Gaze (Standard, 8 XP)} ℎfill \\
\emph{Scene; Zone; No} \\
\textbf{Tags:} \texttt{[STORM]}, \texttt{[FEAR]}, \texttt{[FORCE]} \\
\textbf{Materials:} A mirror polished under a new moon, or a piece of sky‑stone (meteorite fragment). \\
\textbf{Effect:} Call down a brief flash of lightning or a clap of thunder. All who see or hear it within the zone must test \textit{Spirit} (DV 4) or be Stunned for one exchange. Cannot be used indoors or underground. The lightning always strikes the highest point in the zone, which may be inconvenient. \\
\textbf{Push It:} The storm lingers for the scene. All ranged attacks suffer –1 die, and any fire in the zone is extinguished. Mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Monsoon’s Due (High, 14 XP)} ℎfill \\
\emph{Extended; Zone; No} \\
\textbf{Tags:} \texttt{[WEATHER]}, \texttt{[SACRIFICE]}, \texttt{[FATE]}, \texttt{[DEBT]} \\
\textbf{Materials:} A visible scar on your own flesh (cut during the rite) and a handful of earth from the place you wish to affect. \\
\textbf{Effect:} Once per season, you may pray for rain (or clear skies) over a one‑mile area. The weather shifts within an hour and lasts for a full day. If you pray while standing under open sky, you may also ask for a specific omen — a double rainbow, a lightning strike on a named target, or a wind that carries a message to a distant friend. The cost is a permanent scar; Kuva will remember your debt. \\
\textbf{Push It:} The weather shift is immediate and dramatic (a cloudburst or a sudden clearing). Mark \textbf{3 SB (Spades)} as the sky’s disruption draws attention, and the scar never fully heals. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 7 segments.

⫕section*{Cantors & Cults of Kuva}

\textbf{The Cantor’s Song.} A Cantor of Kuva sings in weather‑tides — melodies that rise with the wind and fall with the rain. Their voice is used to call a breeze for becalmed ships, to quiet a storm for a frightened village, or to sing a warning of an approaching typhoon. Kuva’s Cantors are often found on hilltops, in lighthouse towers, or lashed to the masts of dhows, their songs torn from their throats by the very wind they command.

\textbf{The Cult of the Unclouded Eye.} Kuva’s cults meet on high places: mountain peaks, clifftops, the roofs of the tallest buildings. The \textit{Unclouded Eye} is a fellowship of sky‑watchers, weather‑readers, and oath‑sworn travellers. A ritual involves each member naming a promise they have kept and a promise they have broken, then throwing ash into the wind. If the wind carries the ash away, the broken promise is forgiven; if it falls back, the debtor must pay before the next new moon. Their creed is carved on a single stone that is never moved: \textit{“The sky does not forgive. It only reschedules.”} Outsiders are welcome to watch, but they must stand with their faces to the wind and speak no word for the duration of the rite.

⫕section*{Witchcraft of Kuva (Lay / Weather)}

A witch who walks with Kuva is a \textit{wind‑reader} or \textit{storm‑caller} — one who listens to the dome. Their magic works through the instruments of weather: a net of knotted cords that catches wind omens, a bowl of water that reflects the next day’s sky, or a brass whistle that mimics the cry of a storm‑bird. Their tools are not hammers; they use tuning forks, kites, and the small iron bells that ring only when the air pressure drops. In the Mountain Satrapies, an Aeler vent‑prior might calibrate a breath‑book, but a witch of Kuva would climb to a peak and taste the wind to know whether the pass will be safe.

\textbf{The Witch’s Tool: The Wind‑Knot Whistle.} A whistle carved from a bone of a bird that died flying into a storm. Once per session, the witch may blow the whistle in any of the eight directions. For the next hour, the wind in that direction favours the party: +1 Speed on travel, +1 die on any roll involving sails, and the first weather‑based hazard is ignored. However, the wind remembers — the next time the party travels the same route, the GM gains 2 SB to spend on a storm complication.

\textbf{A Hedge Gift: Kuva’s Eavesdrop (2 XP).} Once per night, if you sleep under open sky, you may ask one yes/no question about the next day’s weather. The GM answers truthfully, but the question must be whispered to the wind. On a roll of a 1, the wind answers but the Hollow also hears. The Hollow will remember the question and may send an Oath‑Keeper to ask it back.

⫕section*{Kuva’s Corruption Manifestations}

\begin{tblr}{
  colspec = {Q[c,1.8cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Level & Benefit & Cost / Quirk \\
1 & \textbf{Weather‑Sense:} +1 die to predict rain, wind, or fog. & \textbf{Thin‑Blooded:} –1 die to resist cold; you shiver when the sky is clear. \\
2 & \textbf{Wind’s Favour:} Once per scene, reroll a failed navigation or sailing roll. & \textbf{Compulsive Gaze:} You must look up before speaking a promise; if you forget, the GM gains 1 SB. \\
3 & \textbf{Storm Heart:} +2 dice to resist fear from storms; your voice carries over wind. & \textbf{Restless:} –1 die to social rolls in buildings without windows. \\
4 & \textbf{Lightning’s Child:} Once per session, redirect a lightning strike to a target of your choice (must be outdoors). & \textbf{Static Brand:} Your hair stands slightly on end when you are angry; –1 die to Deception. \\
5 & \textbf{Monsoon’s Voice:} Once per session, call a brief rain that extinguishes non‑magical fires and gives +1 Position to Stealth. & \textbf{Sky‑Debt:} Every use of this power adds 1 tick to a Hollow Attention timer — the Hollow is watching. \\
6+ & \textbf{Unclouded Eye:} Once per session, see the weather of any place you have visited within the last nine days as if you were there. & \textbf{Dome‑Bound:} You cannot enter a cave or cellar without suffering 2 Fatigue; the sky is your only roof. Your eyes are the colour of a storm cloud, and you are always aware of the exact time until the next sunrise. \\
\end{tblr}

\subsection*{Playstyle Notes}
Kuva rewards those who speak truth under open sky, who respect the weather, and who remember that every promise is a debt owed to the dome. Followers become masters of navigation, weather‑reading, and sky‑witnessed oaths. The corruption progression leads toward a being that can no longer bear a roof over its head — but one that can read the sky’s mood from a hundred miles away. Ideal for guides, sailors, rangers, and anyone who has ever watched a storm roll in and felt a thrill of awe.

\subsection*{Rivalries & Obligations}

\textbf{Major Rivalries:}
\begin{itemize}
  ℹtem \textbf{Khemesh (Abyssal Maw):} Sky versus deep. Kuva’s followers refuse to bargain with the pressure beneath the keel; Khemesh’s servants mock the “empty dome.” Rivalries often play out at sea, where storms and tidal waves compete.
  ℹtem \textbf{The Pale Shepherd (Thresholds):} The Shepherd guides souls across the boundary; Kuva simply watches. They are not enemies, but Kuva’s followers distrust any power that claims to speak for the dead.
\end{itemize}

\textbf{Hard Obligation Triggers:}
\begin{itemize}
  ℹtem \textbf{7+ Segments:} You must stand on a hilltop during a thunderstorm and speak a broken oath aloud. If the lightning spares you, the debt is cleared. If not, the Hollow collects.
  ℹtem \textbf{10+ Segments:} Kuva demands a permanent weather station — you must build a shrine on a peak that the sky can see. Refusal means you lose the ability to predict weather; every travel leg becomes Desperate for navigation.
\end{itemize}

\noindent\textbf{Emphasises}
\begin{itemize}
  ℹtem \textbf{Witness of Oaths:} the sky records every promise; break one, and the wind will carry the news
  ℹtem \textbf{Weather as Narrative:} storms are not obstacles but characters with moods
  ℹtem \textbf{Open Sky as Sanctuary:} roofs are prisons, not shelter
  ℹtem \textbf{The Hollow’s Boundary:} where the sky ends, the Hollow begins
\end{itemize}

% −−− Index entries for Kuva −−−
∈dex{Patrons!Kuva}
∈dex{Kuva}
∈dex{Patrons!Sky!Kuva}
∈dex{Patrons!Storms!Kuva}
∈dex{Obligation!Kuva}
∈dex{Rites!Kuva}
∈dex{Patron's Gift!Kuva}