% --- Patron: Mab, Queen of Courts (Glamour & Bargain) ---
\subsection{Mab, Queen of Courts (Glamour \& Bargain)}

\textit{Lore.} Mab rules not from throne or blade, but from dance and debt. She is the smile that binds, the jest that ensnares, the hostess who makes guests complicit in her game. To speak in her Court is to pay; to receive her token is to owe.

Mab is an Archfae – one of the great powers of the Gloaming Courts, neither wholly cruel nor kind, but always bound by her own rules. She cannot lie, but she can mislead so beautifully that truth becomes irrelevant. She cannot break an oath, but she can interpret it until it screams. Her followers thrive on charm, wit, and story, spreading webs of bargains too subtle to escape. The Cantor's Path sings her name most sweetly, for every verse carries a price.

Among the peoples of Fate's Edge, Mab is the patron of courtiers, actors, and those who live by their wits. In Aelinnel, she is the \textit{Unbounded Guest}, whose arrival at a feast is marked by laughter and shattered glass – and a new debt for the host. In Vilikari salons, she is the \textit{Gilded Tongue}, invoked before every negotiation. Wherever masks are worn and words have double meanings, Mab watches.

\begin{quote}
``Every laugh is a promise. Every promise is a debt. Every debt belongs to Mab.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Fae Glamour:} illusions, enchantments, the art of seeming
  \item \textbf{Social Bargains:} oaths, promises, debt as social glue
  \item \textbf{Courtly Intrigue:} etiquette, wordplay, the game of masks
  \item \textbf{Trickster Arts:} mischief, cleverness, the twist in the tale
  \item \textbf{Archfae Sovereignty:} Mab’s word is law in her presence, but only if you agreed to play
  \item \textbf{The Laughing Mask} — Mab, Queen of Courts. \textit{The jest that binds, the mask that becomes the face, the bargain you laughed at and now cannot forget.}
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Mab does not bind a familiar of raw power. Their \textbf{Thiasos} is a small creature of glamour and wit: a talking cricket that gives advice in riddles, a cat with mismatched eyes that appears at masquerades, a small fox that steals shiny baubles and leaves cryptic notes, a white jay that mimics voices with uncanny accuracy, or a spider that weaves its web in the shape of a courtly dance. This creature requires no food, but it must witness a promise or a performance each day. If neglected for a week, it becomes \textit{Compromised} (it pulls pranks at the worst moments – tugging a sleeve during a lie, knocking over a wine glass – imposing –1 die on all Deception and Performance rolls for its bearer until it next witnesses a genuine oath).

The \textbf{Codex} of Mab is never a book of straightforward text. It may be a deck of playing cards whose suits shift when no one is looking, a mirror that reflects only the backs of words, a set of silk handkerchiefs knotted into a puzzle, or a scroll that must be read while dancing. Because Mab’s truths are written in misdirection, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week re‑ordering the cards (or re‑knotting the handkerchiefs) while recounting a story from memory, or the patterns become scrambled (treat as Compromised until re‑woven).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Mab may touch a held item (weapon, tool, or garment) and whisper a jest. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Deception} or \textit{Performance} rolls, and the first time each scene the bearer would be caught in a lie, they may roll Deception again (the glamour twists the truth). After the scene, the user marks \textbf{+1 Obligation} as the memory of a laugh they cannot place follows them.

\subsubsection*{Rites of Mab}

\paragraph{Rite of the Trickster’s Bargain (Low, 4 XP)} \hfill \\
\emph{Scene; Near; Standard Push} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[BOND]}, \texttt{[DEBT]}, \texttt{[FEAR]} \\
\textbf{Materials:} A token freely given (flower, coin, ribbon). \\
\textbf{Effect:} Offer a fae bargain. The target may accept (both gain \textbf{+1 die} to actions that uphold the terms) or refuse (target marks 1 Fatigue from social pressure). Create a 4‑segment \textit{Bargain’s Edge} clock. \\
\textbf{Push It:} If the target later betrays the bargain, they suffer Harm 1 (Stress) and start an \textit{Oathbreaker [4]} clock. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of Courtly Guise (Low, 4 XP)} \hfill \\
\emph{Scene; Self; Standard Push} \\
\textbf{Tags:} \texttt{[GLAMOUR]}, \texttt{[SOCIAL]}, \texttt{[MASK]} \\
\textbf{Materials:} A sprig of hawthorn or a silver thread worn visibly. \\
\textbf{Effect:} Subtle glamour makes you appear as the expected rank or guest. Gain \textbf{+1 die} to social rolls in refined settings. \\
\textbf{Push It:} Mask one personal tell (an accent, a scar, a nervous habit); the first probing question about it generates \textbf{1 SB (Hearts)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Price Agreed (Standard, 8 XP)} \hfill \\
\emph{Scene; Near; Standard Push} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[BOND]}, \texttt{[DEBT]}, \texttt{[TRUTH]} \\
\textbf{Materials:} Equal tokens exchanged ritually (two rings, two signed notes, two locks of hair). \\
\textbf{Effect:} Bind a formal bargain. While the bargain is honoured, both parties gain \textbf{+1 Effect} when cooperating. Breach forces \textbf{1 SB (Hearts/Diamonds)} and a reputation penalty (–1 die to social rolls involving honour for one week). Create an 8‑segment \textit{Bargain Binding} clock. \\
\textbf{Push It:} Add a minor boon to one party (+1 die once). If that party breaches, they suffer an additional 1 SB and you mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Sovereign Glamour (High, 13 XP)} \hfill \\
\emph{Ritual; Zone; High Push} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[GLAMOUR]}, \texttt{[COMMAND]}, \texttt{[WARD]}, \texttt{[FEAR]} \\
\textbf{Materials:} A circle of silk or green felt, courtly music, and a drop of honey (for sweetness) and vinegar (for the sting). \\
\textbf{Effect:} Establish a fae court in the zone. For the scene:
\begin{itemize}
  \item Allies gain \textbf{+1 die} to social rolls; blunt threats (Intimidation, crude Force) suffer –1 die.
  \item Once per scene, you may strip an illusion or reveal a hidden motive as a free action (target tests Resolve DV 3 to resist).
  \item Violating a promise made within the zone triggers \textbf{2 SB} and a visible glamour mark (a blush, a stutter) that lasts a day.
\end{itemize}
Start a 10‑segment \textit{Court Dominion} clock. \\
\textbf{Push It:} Impose a temporary \textit{Court Law} (e.g., “no steel,” “speak only truth,” “every third word must be a compliment”). Violators suffer 2 SB. Mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Gilded Jest (High, 14 XP)} \hfill \\
\emph{Extended; Large Zone; High Push} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[GLAMOUR]}, \texttt{[MASK]}, \texttt{[BOND]}, \texttt{[TRUTH]} \\
\textbf{Materials:} Masks for all participants (hand‑made, never bought), enchanted music, and a cup of wine that has been blessed (or cursed) with a whispered lie. \\
\textbf{Effect:} Create a lasting glamoured court – a masquerade that persists for the duration of the event (hours or an entire evening). All social magic (Sway, Performance, Deception) gains \textbf{+1 Effect}. Outsiders who enter suffer –1 die to perceive truths, read emotions, or see through disguises. Any promise made while masked is magically witnessed; breaking it inflicts \textbf{2 SB} and the mask of the oath‑breaker cracks (visible to all). \\
Start a \textit{Glamour Maintenance [8]} clock; when it fills, the masquerade ends abruptly, and all participants remember everything they said – and forgot. \\
\textbf{Push It:} The masquerade persists between scenes for the rest of the session. Mark \textbf{+3 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 8 segments.

\subsubsection*{Cantors \& Cults of Mab}

\textbf{The Cantor's Song.} A Cantor of Mab sings in \textit{harmonies of the half‑truth} – melodies that shift key without warning, leaving the listener uncertain of what was heard. Their voice is used to seal a bargain with a sung clause, to make a crowd forget a social faux pas, or to weave a glamour that lasts as long as the music plays. Mab's Cantors are prized at every court where masks are worn, and feared by those who know that a beautiful tune can hide a poisoned promise.

\textbf{The Cult of the Gilded Jest.} Mab's cults meet in places where masks are natural: masquerade balls, carnival tents, and the back rooms of theatres. The Gilded Jest is less a religion and more a conspiracy of wits – actors, gamblers, and lawyers who have learned that every exchange is a performance. A ritual of the Jest involves each member writing a true secret on a playing card, then exchanging cards with another member. They must then speak the secret they received as if it were their own – and the group votes on whether they told it “true enough.” Those who fail must drink a cup of “tea of forgotten clauses” (a bitter brew of nettle and fae‑touched honey) and lose one minor memory (a face, a password, a small debt). Their creed is printed on a card that is passed, never kept: \textit{"The best chains are made of laughter. The strongest oaths are sworn in jest."} Outsiders are welcome as “audience,” but those who laugh too loudly may find themselves on stage – and the stage has its own rules.

\subsubsection*{Witchcraft of Mab (Lay / Glamour)}

A witch who walks with Mab is a \textit{glamour‑weaver} or \textit{toy‑maker} – one who crafts objects that are never quite what they seem. Their magic works through misdirection: a doll that laughs when a promise is broken, a mirror that shows a flattering reflection (and nothing else), or a mask that lets the wearer speak any lie as truth. Their tools are never hammers; they use needles, masks, and the subtle art of a well‑placed reflection. In Aeler, a True‑Mason might carve a cornerstone that makes everyone forget the building’s weak point, but a witch of Mab would instead sew a glamour seam into the cornerstone’s shadow.

\textbf{The Witch’s Tool: The Mask of Shifting Faces.} A simple half‑mask of polished white wood, its eye holes rimmed with silver. Once per session, the witch may don the mask and change their apparent appearance – face, voice, and even perceived gender – for one scene. This grants \textbf{+2 dice} to \textit{Deception} or \textit{Performance} rolls to pass as someone else, and anyone who tries to see through the disguise must test Resolve (DV 4). The mask cannot mimic a specific person unless the witch has a token of that person (a hair, a signature). After use, the mask’s eyes weep a single drop of silver; the witch may collect it as a component for glamour rites.

\textbf{A Hedge Gift: Mask of the Moment (2 XP).} Once per scene, the witch may place a hand over their own face and draw it away, revealing a different expression (or a different face entirely, briefly). This grants \textbf{+1 die} to a single \textit{Deception} or \textit{Performance} roll, as the target’s expectations are momentarily overturned. The illusion lasts only for that exchange.

\subsubsection*{Mab's Corruption Manifestations}

\begin{tblr}{
  colspec = {Q[c,1.8cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Level & Benefit & Cost / Quirk \\
1 & \textbf{Glamour’s Touch:} +1 die to Deception/Performance with stories. & \textbf{Truthbound:} Cannot speak plain falsehood; only mislead. \\
2 & \textbf{Fairy Step:} Once per scene, flicker Near as if teleporting. & \textbf{Iron Weakness:} 1 Fatigue if touched by cold iron. \\
3 & \textbf{Trickster’s Delight:} Spend 1 Boon to twist a Complication advantageously. & \textbf{Compulsive Jest:} Must play a trick each session or mark 1 SB. \\
4 & \textbf{Hospitality’s Gift:} Allies sharing food/drink gain +1 die to Resolve. & \textbf{Hospitality Bound:} Harming guests costs +2 Obligation. \\
5 & \textbf{Fae Sight:} See veils/glamours; +2 dice to Notice. & \textbf{Truth Debt:} Must accept “fair” trades or suffer. \\
6+ & \textbf{Twilight Crown:} Once/session, declare an Oath; you gain +2 dice to all rolls that fulfil it, and enemies who swore the same oath suffer –1 die to resist its terms. & \textbf{Oathbound:} Breaking this oath inflicts Harm 2 (Stress) and starts an \textit{Iron Debt [6]} clock that can only be cleared by a quest. Your shadow is now forked – a visible sign that Mab has marked you. \\
\end{tblr}

\subsection*{Playstyle Notes}
Mab excels in social manipulation and fae‑style bargaining. Followers become masters of subtle magic, binding others through oaths and glamour. The corruption progression leads toward becoming fae‑touched, gaining power over truth and perception at the cost of mortal constraints. Intrusions often take the form of a forgotten bargain resurfacing, a guest demanding hospitality you cannot refuse, or a jest that turns cruel. Ideal for players who enjoy social intrigue, wordplay, and characters who prefer cleverness over brute force.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Archfae Sovereignty:} Mab’s word is law in her presence, but only if you play her game
  \item \textbf{Glamour as Social Weapon:} illusions that bind, masks that deceive
  \item \textbf{Bargains and Debts:} every gift is a hook, every promise a leash
  \item \textbf{Courtly Intrigue:} etiquette, wordplay, the dance of masks
  \item \textbf{Trickster Wisdom:} the truth is what you can make others believe
\end{itemize}