% ============================================================================
% PATRON ENTRY: MOR'IRAATH --- THE DESTROYER (ANNIHILATION & INFERNO)
% ============================================================================

\subsection{Mor'iraath --- The Destroyer (Annihilation \& Inferno)}

\subsubsection*{Lore}
Mor'iraath is the demon of absolute destruction---not the slow decay of rot, nor the patient erosion of water, but the joyful, all-consuming fire that reduces everything to ash and memory. He delights not in the process of ruination, but in the perfect moment of combustion, when something that was becomes something that never will be again.

His followers become agents of finality, wielding flame that leaves no trace, no remains, and no possibility of reconstruction. To serve him is to embrace the beauty of endings and the power that comes from making things never have been. Where Malachai corrupts and Khemesh crushes, Mor'iraath \textit{ignites}.

Among the peoples of Fate's Edge, Mor'iraath is known by fearful names. In Zakov, he is the \textit{Sea-Scourge}, invoked by pirates who burn their prizes to the waterline. In Aeler, he is the \textit{Deep Forge} that consumes ore and returns only slag. In Linn, he is the \textit{Tide-That-Eats-The-Shore}, a slow, patient consumption that reshapes coastlines. His followers are rarely beloved, but they are essential---for without destruction, there is no room for creation.

\begin{quote}
``Every treasure is fuel. Every pact is tinder. Ash is the only inheritance.''
\end{quote}

\subsubsection*{Domain Focus}
\begin{itemize}
  \item \textbf{Fiery Annihilation:} fire that consumes utterly, leaving only ash
  \item \textbf{Absolute Endings:} no trace, no memory, no return
  \item \textbf{Cleansing Flame:} burning away corruption, weakness, and existence
  \item \textbf{Inferno as Art:} the beauty of a perfect, final combustion
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}

A Runekeeper sworn to Mor'iraath does not bind a familiar of hearth and home---they bind a companion of ash and ember. Their \textbf{Thiasos} is a small creature touched by finality: a blackened salamander that sleeps in hot coals, a fire-scorpion whose sting leaves a cauterized wound, a soot-colored moth that leaves ash trails on whatever it touches, or a tiny phoenix fledgling that cannot yet rise from its own ashes (and may one day burn its keeper). This creature requires no food, but it must be fed a small amount of combustible material (charcoal, oil, or a burnt offering) each day. If neglected for a week, it becomes \textit{Compromised} (it smolders constantly, emitting smoke that imposes --1 die on all Stealth rolls and attracts the attention of fire-wardens).

The \textbf{Codex} of Mor'iraath is never a book of paper and ink. It may be a slab of fused glass etched with runes of ending, a set of iron plates that must be heated before the Rites become visible, or a scroll of asbestos that burns without being consumed. Because Mor'iraath's truths are written in ash, the Codex requires upkeep: the Runekeeper must spend an hour each week burning a small offering (a scrap of vellum, a lock of hair) and letting the smoke imprint the Rites anew, or the symbols fade (treat as Compromised until re-burned).

\subsubsection*{Patron's Gift (Imbuement)}

Once per scene as an action, a Runekeeper of Mor'iraath may touch a held item (weapon, tool, or garment) and whisper a word of unmaking. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Melee} or \textit{Ranged} attacks, and the first time each scene the bearer would reduce an enemy to 0 Harm, they may choose to utterly destroy that enemy (no remains, no regeneration, no trace). After the scene, the user marks \textbf{+1 Obligation} to Mor'iraath.

\subsubsection*{Rites of Mor'iraath}

\paragraph{Rite of the Perfect Ending} (Basic, 4 XP) \hfill \\
\emph{Instant; Touch; No.} \\
\textbf{Tags:} \texttt{[RADIANT]}, \texttt{[CONSUME]}, \texttt{[ERASURE]} \\
\textbf{Materials:} The final component of something (last brick, final word, closing signature), or a pinch of ash from a previous fire. \\
\textbf{Effect:} Cause one small object to be utterly consumed by flame---no remains, no trace, no possibility of reconstruction. Gain \textbf{+1 die} to rolls involving finality or absolute conclusions. \\
\textbf{Push It:} The ending creates a beneficial void---gain 1 Boon as space opens for something else; mark \textbf{1 SB (Spades)} as reality adjusts to the absence. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Instant.

\paragraph{Rite of the Unmaking Gaze} (Basic, 5 XP) \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[RADIANT]}, \texttt{[OMEN]} \\
\textbf{Materials:} A lens of obsidian or fire-scarred glass. \\
\textbf{Effect:} \textbf{+2 dice} to identify the precise weakness that will cause complete collapse. See the ``breaking point'' in any structure, plan, or construct---the spot where a single spark would bring it all down. Your next destructive action against the identified target gains \textbf{+1 Effect}. \\
\textbf{Push It:} The gaze becomes prophetic---see three possible endings for one target; choose which occurs, but mark \textbf{1 SB (Diamonds)} as fate resists your vision. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Cleansing Flame} (Advanced, 8 XP) \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[RADIANT]}, \texttt{[AREA]}, \texttt{[ERASURE]}, \texttt{[BARRIER]} \\
\textbf{Materials:} Fuel that was once part of what you intend to destroy (a beam from the building, a lock of hair from the person). \\
\textbf{Effect:} Create a fire that burns not just substance but essence---what is consumed by this flame never existed. The zone becomes a \textit{pyre zone}; anything entering risks being marked for erasure. Enemies suffer \textbf{--2 dice} to actions that would preserve or continue anything within the zone. \\
\textbf{Push It:} The flame becomes selective---choose what is unmade and what remains; mark \textbf{+1 Obligation} as you assume judgment over existence itself. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Absolute Negation} (Advanced, 9 XP) \hfill \\
\emph{Instant; Near; No.} \\
\textbf{Tags:} \texttt{[RADIANT]}, \texttt{[BANISH]}, \texttt{[FATE]}, \texttt{[FEAR]} \\
\textbf{Materials:} The complete destruction of a representation of the target (a lock for a door, a written name for a person, a fragment for a whole), burned to ash. \\
\textbf{Effect:} The target suffers \textbf{--2 dice} to resist any effect that would cause their non-existence. If they are reduced to Harm 0 by your actions, they are utterly consumed by infernal flame---\texttt{[BANISH]}ed from all realities, never were, never could be. \\
\textbf{Push It:} The negation becomes retroactive---the target never existed in your presence; mark \textbf{2 SB (Hearts)} as the timeline rewrites around their absence. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Instant.

\paragraph{Rite of the Unraveling Inferno} (Master, 13 XP) \hfill \\
\emph{Scene; Touch; No.} \\
\textbf{Tags:} \texttt{[RADIANT]}, \texttt{[AREA]}, \texttt{[FATE]}, \texttt{[DEATH]} \\
\textbf{Materials:} The foundational element of what you wish to unmake (keystone, cornerstone, source code, a written covenant). \\
\textbf{Effect:} Your touch ignites a cascading dissolution. The target and all things fundamentally connected to it begin immediate combustion, crumbling to ash as if they had never been. Test DV = Complexity.
\begin{itemize}
  \item Success: Target and connections cease to exist.
  \item Partial: Significant damage, accelerated decay, but some remnants remain.
  \item Miss: Backlash; the fire turns on you---mark 2 SB and suffer Harm 1 (Burn).
\end{itemize}
\textbf{Push It:} The unraveling becomes contagious---similar things in Near range also ignite; mark \textbf{+2 Obligation}. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Final Pyre} (Master, 14 XP) \hfill \\
\emph{Extended; Zone; No.} \\
\textbf{Tags:} \texttt{[RADIANT]}, \texttt{[AREA]}, \texttt{[ERASURE]}, \texttt{[PRESSURE]} \\
\textbf{Materials:} Complete dedication of the space to Mor'iraath---every flammable object arranged as a funeral offering, and a drop of your own blood to seal the pact. \\
\textbf{Effect:} The zone becomes a pocket of absolute combustion. Anything that enters is set ablaze, suffering Harm 1 each round (ignores Armor). At the end of each round, they must test Spirit+Resolve (DV 5) or be consumed entirely (cease to exist). The space itself becomes a \textit{pyre scar}---it remains hot to the touch for a year and a day, and nothing will grow there. \\
\textbf{Push It:} The pyre becomes permanent and expands at 1 segment per scene; you suffer Harm 2 (Burn) and mark \textbf{+3 Obligation} as you become partially unmade yourself, your left hand now permanently blackened and smoking. Cost: Mark +1 Obligation (total +4) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Permanent.

\subsubsection*{Cantors \& Cults of Mor'iraath}

\textbf{The Cantor's Song.} A Cantor of Mor'iraath sings in \textit{notes of cessation}---melodies that start strong and fade to crackling silence, like a fire burning out. Their voice is used to silence a battlefield after a massacre, to erase a name from a town's records, or to sing a dying person gently into nothingness rather than an afterlife. Mor'iraath's Cantors are terrifying not because they are cruel, but because they offer a perfect, peaceful, absolute end---and some people accept.

\textbf{The Cult of the Final Breath.} Mor'iraath's cults gather in places where things end: funeral pyres, abandoned ruins, collapsed mines, and the cells of the condemned. The Final Breath is a fellowship of nihilists, assassins of memory, and those who have lost everything and now embrace the cleansing flame. A ritual involves each member writing down something they wish had never existed---a memory, a person, a part of themselves---and then burning the paper in a brazier that gives no heat. They then drink a cup of ash suspended in water; the ash is from a fire that consumed something they once loved. Their creed is whispered, never written: \textit{``What is not cannot suffer. What is not cannot be hurt. We give the gift of never-having-been.''} The most devoted tattoo a single ember on the inside of their lower lip; the tattoo smokes when they speak a lie. Outsiders are not welcome; they are targets, or potential converts who have already lost too much to fear the flame.

\subsubsection*{Witchcraft of Mor'iraath (Lay / Unmaking)}

A witch who walks with Mor'iraath is a \textit{pyre-smith} or \textit{ash-weaver}---one who crafts through the art of burning. Their magic works through the materials of finality: a stamp that brands a ``flaw'' into a threshold, a quill that writes in ink made from burnt secrets, or a forge that tempers a blade that leaves no wound (only a cauterized absence). In Aeler, a True-Mason might carve a keystone that makes the tunnel behind it never have existed, but a witch of Mor'iraath would instead press a hot iron sigil into the keystone, and the tunnel would collapse into ash.

\textbf{The Witch's Tool: The Cinder-Stamp.} An iron rod with a sigil of ending carved into its face, heated until it glows dull red. Once per session, the witch may press the stamp against a threshold (a door, a bridge, a city gate) while speaking a single word---the name of something that should never have existed. For the next scene, anyone who crosses that threshold and fails a Spirit+Resolve test (DV 3) will forget the last minute of their life---their memory is burned away, not suppressed. The stamp can also be used to ``unwrite'' a single word from a document; the word becomes a smudge of ash, and the sentence restructures itself around the absence.

\textbf{A Hedge Gift: Absence Knife (Basic, 2 XP).} Once per scene, the witch may draw a finger (as if holding a knife) across a surface---a lock, a rope, a clasp---and speak a word of Mor'iraath's name. The object does not break; it simply ceases to be an obstacle. The lock is no longer locked (there is no lock). The rope is no longer a rope. The clasp is as if it never had a function. This effect lasts for one exchange, after which reality ``repairs'' itself, but the moment of absence is enough.

\subsubsection*{Corruption of Mor'iraath}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Ending's Sight:} +1 die to Notice when something is approaching its natural conclusion or final form. & \textbf{Completion Anxiety:} Must resist urge to ``finish'' things prematurely; --1 die to rolls involving patience or process. \\
2 & \textbf{Perfect Strike:} Once/scene, your destructive action gains +2 dice and ignores all resistance to ending/ceasing. & \textbf{Destruction Proximity:} Suffer 1 Fatigue when near things of personal value; their existence pains you. \\
3 & \textbf{Unmaking Touch:} +2 dice to rolls involving absolute destruction---no remains, no possibility of return. & \textbf{Entropy Aura:} Objects of value near you show signs of decay/premature aging; allies suffer --1 die to rolls involving cherished possessions. \\
4 & \textbf{Final Judgment:} Once/session, declare one thing in Near range must make Spirit+Resolve (DV 4) or cease to exist entirely. & \textbf{Existential Hunger:} --1 die to rolls involving creation, preservation, or continuation; you hunger to end what others would sustain. \\
5 & \textbf{Principle of Negation:} Once/session, unmake an abstract concept (hope, fear, memory) from a willing target. & \textbf{Conceptual Erosion:} Your presence causes ideas and beliefs to weaken; --1 die to rolls requiring stable reality or consistent identity. \\
6+ & \textbf{Avatar of Unbeing:} Once/session, become the living embodiment of annihilation---anything you touch burns to ash and ceases to exist for one scene. & \textbf{Void Incarnate:} Mark +3 Obligation when using; risk becoming a permanent fire that must constantly consume to maintain existence, eventually burning even yourself. \\
\end{tblr}

\subsection*{Playstyle Notes}
Mor'iraath is the demon of fiery destruction---the joy of a perfect, all-consuming blaze. His followers gain power over absolute endings, the ability to erase things from existence, and the terrible beauty of ash. The cost is the loss of patience, connection, and eventually the ability to value anything that lasts. Intrusions often take the form of a cherished object crumbling to ash, a fire that cannot be extinguished, or a demand to end something that you love. Ideal for players who enjoy nihilistic themes, the power of finality, and characters who have decided that if nothing matters, then nothing deserves to remain.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Fiery Annihilation:} fire that leaves nothing behind
  \item \textbf{Absolute Endings:} no trace, no memory, no return
  \item \textbf{Cleansing Flame:} burning away corruption, weakness, and existence itself
  \item \textbf{Inferno as Art:} the beauty of a perfect combustion
  \item \textbf{Ash as Legacy:} all that remains of empires, lovers, and gods
\end{itemize}

% --- Monastic Tradition: Ash-Fist (Mor'iraath) ---
\begin{tcolorbox}[colback=black!5,colframe=red!70!black,title=\textbf{Ash-Fist (Mor'iraath)},breakable]
  \emph{Every treasure is fuel. Every pact is tinder.}

  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Body 3+, Spirit 2+, Rite of the Perfect Ending or Cantor of Mor'iraath.

  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} Ash cannot be held to account. Monks of the Ash-Fist gain \textbf{+1 die} to resist any attempt to compel them through magical debt, oath, or contract---but each such resistance leaves a faint smell of smoke, and the GM may gain 1 SB (the universe remembers a debt that was burned rather than paid).

  \vspace{2mm}
  \noindent\textbf{Techniques (purchased as Talents):}

  \begin{tblr}{
    colspec = {X[0.7,l] X[1.8,l] X[1.1,l] X[1.4,l]},
    rowhead = 1,
    row{odd}={bg=gray!10},
    row{1}={bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Ending Touch} (Basic, 5 XP) & Declare before rolling. On a hit, the target leaves no remains (cannot be resurrected, tracked, or remembered by magical means). On a miss, you take Harm 1 (Burn). & 1 Boon & Body 3 \\
  \textbf{Unmaking Gaze} (Basic, 4 XP) & +2 dice to identify the precise weakness that will cause complete collapse. Your next destructive action against that target gains +1 Effect. & 1 Fatigue & Wits 2 \\
  \textbf{Cleansing Flame} (Basic, 6 XP) & As an action, create a small fire that burns essence. Any flammable object you touch can be utterly consumed---no trace. & 1 Boon & Spirit 2 \\
  \textbf{Absolute Negation} (Advanced, 9 XP) & When you reduce an enemy to 0 Harm, you may declare them utterly annihilated (never existed). & Free & Tier II \\
  \textbf{Ash Memory} (Basic, 6 XP) & Once per session, you may ``remember'' a fact that was erased (as if it had never been lost)---but you also forget one minor memory of your own (GM's choice). & Free & Spirit 3 \\
  \end{tblr}

  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Void Incarnate.} Once per session, become living annihilation for one scene: anything you touch (including abstract debts or contracts) can be erased from existence. You may delete one ongoing magical obligation, curse, or debt entirely---but you permanently lose a cherished memory. After the scene, you gain a permanent \textbf{Debt Scar} (a hand that is always slightly smoking). You may ignore one debt or obligation that was recorded under your old identity, but you will also forget a small part of who you were.
\end{tcolorbox}