% --- Patron: Nidhoggr, the World-Worm (Dreaming Antiquity) ---
\subsection{Nidhoggr, the World-Worm (Dreaming Antiquity)}

\textit{Lore.} Beneath stone and sleep coils Nidhoggr, who gnaws at the roots of time. He does not speak quickly; he dreams in centuries. To press your ear to the earth is to risk drowning in the silence of aeons. Yet for those who endure, he whispers truths long buried, memories fossilized in stone, and the slow inevitability of cycles unbroken. His followers walk in twilight between dream and ruin, bearing the weight of all that has been.

Nidhoggr is the worm who knows the metaphorical bones of the earth – the secret history hidden in every mountain, the forgotten temple beneath the marketplace, the name of the king whose tomb was erased from every chronicle. He is the patron of archaeologists, relic hunters, and those who understand that the past is never truly dead. It is merely sleeping.

Among the peoples of Fate's Edge, Nidhoggr is called by many names. In Aeler, he is the \textit{World‑Worm}, whose dreaming coils are said to cause earthquakes. In the Mistlands, he is the \textit{Fog‑Memory}, the presence that makes the mist remember what the living forget. In Ykrul, he is the \textit{Bone‑Reader}, whose whispers guide shamans to ancient burial grounds. Wherever the earth holds secrets, Nidhoggr dreams.

\begin{quote}
``Press your ear to the earth and wait. If it remembers you, it will answer.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Ancient Knowledge:} forgotten histories, buried truths, ancestral memory
  \item \textbf{Stone Communion:} speaking with earth, reading geological memory, stone lore
  \item \textbf{Temporal Perception:} seeing echoes of the past, prophetic dreams, slow time
  \item \textbf{Cyclical Power:} seasonal magic, ritual timing, patience as virtue
  \item \textbf{Relic Hunter:} unearthing the past, preserving what was lost
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Nidhoggr does not bind a familiar of quick flesh. Their \textbf{Thiasos} is a small creature of deep time and patience: a blind salamander that lives in caves and never ages, a fossilised beetle that occasionally twitches its legs, a snail whose shell records the layers of years, or a small burrowing mole whose fur is always dusted with ancient soil. This creature requires no food, but it must spend at least an hour each day in contact with stone or earth (a cave, a grave, a masonry wall). If neglected for a week, it becomes \textit{Compromised} (it grows lethargic and its shell cracks, imposing –1 die on all Survival or Lore rolls involving ancient sites for its bearer until it is returned to stone).

The \textbf{Codex} of Nidhoggr is never a fresh book. It may be a fossilised slab with impressions that change under moonlight, a bundle of slate tablets etched with runes that deepen with age, or a scroll of petrified leather that unrolls only when buried in soil. Because Nidhoggr’s truths are written in the language of deep time, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week burying it in earth (or placing it against ancient stone) and listening for the whisper that refreshes the Rites, or the symbols fade (treat as Compromised until re‑interred).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Nidhoggr may touch a held item (weapon, tool, or archaeological find) and whisper a word of deep memory. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Lore} or \textit{Investigation} rolls when examining ancient objects or sites, and the first time each scene the bearer would fail to notice a trap or hidden feature in a ruin, they may reroll the check (the stone whispers a warning). After the scene, the user marks \textbf{+1 Obligation} as the weight of ages presses on their mind.

\subsubsection*{Rites of Nidhoggr}

\paragraph{Rite of the Whispering Stone (Low, 5 XP)} \hfill \\
\emph{Scene/Instant; Touch/Self; Yes} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[DREAM]}, \texttt{[TIME]} \\
\textbf{Materials:} Dust from weathered stone or an ancient artifact; or water poured over ancient stone, consumed ritually. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Stone's Whisper:} Stone reveals one hidden fact about its history. Gain \textbf{+1 die} to Lore about this object's origin and purpose.
  \item \textbf{Dreaming Deep:} Enter a brief trance to witness echoes of local history. Ask one specific question about past events here.
\end{itemize}
\textbf{Push It:} (Stone) The vision includes sensory details from the past; mark \textbf{+1 Obligation}. (Dream) The vision reveals additional context or hidden connections; mark \textbf{1 SB (Clubs)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Stone-Sleeper's Murmur (Standard, 8 XP)} \hfill \\
\emph{Scene; Zone; Standard Push} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[AREA]}, \texttt{[TIME]} \\
\textbf{Materials:} An ear pressed to bedrock or an ancient structure. \\
\textbf{Effect:} Read accumulated memories in stone. Ask up to 3 questions about significant events witnessed here. Creates a 6‑segment \textit{Memory Resonance} clock. \\
\textbf{Push It:} One memory unfolds with perfect clarity; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of Awakened Chronicle (Standard, 9 XP)} \hfill \\
\emph{Ritual; Zone; Standard Push} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[AREA]}, \texttt{[TIME]}, \texttt{[BOOST]} \\
\textbf{Materials:} A chalk spiral, four local stones, an hour of meditation. \\
\textbf{Effect:} The area replays ghostly echoes of a chosen past event. All witnesses gain \textbf{+2 dice} to Investigate/Recall related checks. Creates an 8‑segment \textit{Echo Stability} clock. \\
\textbf{Push It:} View events from multiple perspectives; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} Extended ritual; mark \textbf{+1 Obligation}.

\paragraph{Rite of the World-Worm's Dream (High, 13 XP)} \hfill \\
\emph{Extended; Zone; High Push} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[DREAM]}, \texttt{[TIME]}, \texttt{[PROPHECY]} \\
\textbf{Materials:} A circle of ancient stones, a night of vigil under stars. \\
\textbf{Effect:} Access deep geological memory. Ask 3 questions about ancient history or receive a prophetic dream. Creates a 10‑segment \textit{Deep Dream} clock. \\
\textbf{Push It:} Gain temporary skill mastery (+3 dice to one historical action); mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Unburied Past (High, 14 XP)} \hfill \\
\emph{Extended; Large Zone; High Push} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[AREA]}, \texttt{[TIME]}, \texttt{[PRESSURE]} \\
\textbf{Materials:} Stones from three ancient sites, the blood of a history‑keeper (a scholar, a chronicler, or a witness). \\
\textbf{Effect:} Overlay a past reality onto the present location. Ruins partially reform, ghostly figures manifest. Allies gain \textbf{+2 dice} to Lore/Investigate; any modern technology or recent construction suffers \textbf{–2 dice} to function within the zone. \\
\textbf{Push It:} The effect lasts until the next lunar cycle; start a \textit{Temporal Instability [8]} clock. When it fills, the past and present become dangerously entangled – a location may become inaccessible or a relic may begin to fade. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 8 segments.

% --- Relic Hunter (Archaeologist/Explorer) tcolorbox ---
\begin{tcolorbox}[colback=black!5,colframe=brown!70!black,title=\textbf{Relic Hunter (Seeker of Lost Ages)}]
\emph{The past is not a foreign country – it is a treasure chest buried beneath your feet. Know where to dig, know what to keep, and know when to run.}

\vspace{2mm}
\noindent\textbf{Prerequisites:} Wits 3+, Lore 2+, Rite of the Whispering Stone or Cantor of Nidhoggr.

\vspace{2mm}
\noindent\textbf{Relic Hunter Techniques (choose as Talents):}

\begin{tblr}{
  colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.3,l]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Technique & Effect & Cost & Req. \\
\textbf{Cartographer's Eye} & Once per expedition (journey), you may redraw a safe route past one environmental hazard (trap, unstable ground, patrol). & 1 Boon & Wits 3 \\
\textbf{Stone Tongue} & You can read ancient scripts and petroglyphs as if fluent, even if you have never seen the language before. & Passive & Lore 2 \\
\textbf{Unstable Ground} & When you are in a ruin or underground, you gain +1 die to evasion rolls (dodging traps, collapsing ceilings, etc.). & Passive & Wits 3, Athletics 2 \\
\textbf{Historical Insight} & Once per scene, when you encounter a relic, inscription, or ancient structure, you may ask the GM one yes/no question about its purpose or origin. & Free & Tier II \\
\textbf{Trap Sense} & You are never surprised by a trap. You gain +1 die to Notice to detect traps, and if you fail, you may reroll one die. & Passive & Wits 3, Perception 2 \\
\textbf{Escape Artist} & Once per session, when you would be trapped (collapsed tunnel, sealed door, net), you may declare a hidden exit or a forgotten passage. The GM reveals a way out, but it may lead to another danger. & 1 Boon & Tier III, Athletics 3 \\
\end{tblr}

\vspace{2mm}
\noindent\textbf{Master Technique (Epic Talent, 12 XP):}
\textit{Secrets of the Unburied Past.} Once per session, you may touch a significant relic or ancient site and ask the GM one question about its history. The GM must answer truthfully and in detail (a paragraph or more). The answer may reveal a hidden weakness, a forgotten ritual, or the location of a related treasure. After using this talent, you gain +1 permanent Corruption (the past weighs on you) and mark +1 Obligation to Nidhoggr – the worm expects you to act on what you have learned.
\end{tcolorbox}

\subsubsection*{Cantors \& Cults of Nidhoggr}

\textbf{The Cantor's Song.} A Cantor of Nidhoggr sings in \textit{geological time} – melodies so slow that a single note may stretch across a minute, harmonics that evoke the groan of shifting plates. Their voice is used to lull a city to sleep so that its buried secrets may be heard, or to sing a stone back to its ancient shape. Nidhoggr's Cantors are often found in deep caves, whispering to stalactites that remember the dawn of the world.

\textbf{The Cult of the Sleeping Stone.} Nidhoggr's cults gather in places where time feels heavy: ancient barrows, fossil beds, the roots of mountains. The Sleeping Stone is a fellowship of dreamers, historians, and those who have learned that the world remembers everything – and sometimes forgets on purpose. A ritual of the Cult involves each member pressing their ear to a stone and listening for a full hour in absolute silence. When one of them hears a whisper, they speak it aloud, and the stone's memory is recorded. Their creed is carved into a single erratic boulder that moves about the landscape once a century: \textit{"What is buried is not lost. What is forgotten is only waiting."} Outsiders are welcome to listen, but they must bring a stone from their homeland and leave it behind.

\subsubsection*{Witchcraft of Nidhoggr (Lay / Stone Memory)}

A witch who walks with Nidhoggr is a \textit{fossil‑scribe} or \textit{stone‑dreamer} – one who crafts with the memory of the earth. Their magic works through the materials of deep time: a piece of amber that preserves a forgotten scene, a lens ground from a fossilised shell that magnifies the past, or a bowl carved from a meteorite that holds visions of the sky when the world was young. Their tools are never hammers; they use polished stones, fossil lenses, and the patient pressure of a thumb on a petroglyph.

\textbf{The Witch’s Tool: The Eye of the World‑Worm.} A smooth, oval stone, polished to a mirror sheen, with a spiral fossil embedded at its centre. Once per session, the witch may hold the stone to their eye and look at any ancient structure, fossil, or relic. For the next scene, they see the last significant event that occurred near that object as a ghostly overlay – the mason carving the keystone, the dinosaur stepping into the tar pit, the priest sealing the tomb. The vision is silent but detailed; the GM must describe one useful fact (e.g., “the entrance was hidden behind a tapestry,” “the fossil’s tail points north”). The stone grows warm when it is ready to be used again.

\textbf{A Hedge Gift: Echo of the Deep (2 XP).} Once per scene, the witch may place a hand on a stone structure (a wall, a bridge, a cliff) and speak a single word. The stone “repeats” the last word spoken in its presence by any creature – but the word is always in the voice of the last person who touched it, even if they are long dead. This can be used to identify the inhabitants of a ruin, or to frighten the superstitious.

\subsubsection*{Nidhoggr's Corruption Manifestations}

\begin{tblr}{
  colspec = {Q[c,1.8cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Level & Benefit & Cost / Quirk \\
1 & \textbf{Stone Sense:} +1 die to detect ancient sites, hidden history, or geological features. & \textbf{Temporal Drift:} Speak in archaic patterns; --1 die to modern social interactions. \\
2 & \textbf{Memory Echo:} Once/session, perfectly recall one historical fact as if witnessed. & \textbf{Heavy Presence:} 1 Fatigue in places with no history or recent construction. \\
3 & \textbf{Earth Whisper:} +2 dice to commune with stone, read echoes, or interpret ancient sites. & \textbf{Rooted:} --1 die to actions requiring rapid movement or aerial maneuvers. \\
4 & \textbf{Dream Prophet:} Once/scene, gain +1 die to actions tied to ancient mysteries. & \textbf{Ancestral Haunting:} Visions interrupt sleep; mark 1 SB if rest is disturbed. \\
5 & \textbf{Timeless Gaze:} Once/session, witness distant past events through stone communion. & \textbf{Geological Pace:} --1 die to reactions requiring haste; act with deliberate slowness. \\
6+ & \textbf{World‑Memory:} Become a living archive for one scene (+3 dice to historical knowledge). & \textbf{Time's Weight:} Mark +2 Obligation; risk permanent Harm from temporal overload. \\
\end{tblr}

\subsection*{Playstyle Notes}
Nidhoggr rewards patience, curiosity, and a reverence for the past. Followers excel as archaeologists, relic hunters, and scholars who uncover truths that others would leave buried. The corruption progression leads characters toward becoming living libraries of forgotten lore, but at the cost of connection to the present moment – and sometimes the loss of their own identity in the weight of ages. Intrusions often take the form of prophetic dreams that blur past and present, a relic that begins to speak, or a demand to return a stolen artifact to its proper grave. Ideal for players who love exploration, mystery, and the thrill of unearthing what was never meant to be found.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Ancient Knowledge:} the past is a resource, not a relic
  \item \textbf{Stone Communion:} the earth remembers everything
  \item \textbf{Relic Hunting:} every ruin is a story, every artifact a clue
  \item \textbf{Temporal Patience:} some secrets take centuries to ripen
  \item \textbf{Metaphorical Bones:} history is not dead – it is dreaming
\end{itemize}

% --- Monastic Tradition: Stone Dreamer (Nidhoggr) ---
\begin{tcolorbox}[colback=black!5,colframe=green!70!black,title=\textbf{Stone Dreamer (Nidhoggr)}]
  \emph{Press your ear to the earth and wait.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Wits 3+, Spirit 2+, Rite of the Whispering Stone or Cantor of Nidhoggr.
  
  \vspace{2mm}
  \noindent\textbf{Debt–Resistant Frame.} The World‑Worm dreams in centuries; debts are but a heartbeat. Stone Dreamer monks gain \textbf{+1 die} to resist any magical debt or obligation that would bind them for a limited time – they simply wait it out, and the creditor’s urgency crumbles like old stone. If they succeed, the obligation is delayed by one day per rank of the monk’s Spirit (minimum one day).
  
  \vspace{2mm}
  \noindent\textbf{Techniques:}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.4,l]},
    rowhead = 1, row{odd}={bg=gray!10}, row{1}={bg=gray!30}, hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Fossil Strike} & You may “store” a missed attack. The stored die can be added to a later attack as bonus dice. Maximum storage = Spirit attribute. & Free & Wits 3 \\
  \textbf{Whispering Stone} & Touch a stone to learn one hidden fact about its history. Gain +1 die to Lore about this object’s origin and purpose. & 1 Fatigue & Spirit 2 \\
  \textbf{Stone‑Sleeper's Murmur} & Read accumulated memories in stone. Ask up to 3 questions about significant events witnessed here. Create a 6‑segment Memory Resonance clock. & 1 Boon & Wits 3 \\
  \textbf{Earth Whisper} & +2 dice to commune with stone, read echoes, or interpret ancient sites. & 1 Fatigue & Tier II \\
  \textbf{Rooted Patience} & When you stand still for one full exchange, gain +1 die to your next action (as the earth steadies you). & Passive & Spirit 2 \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{World‑Memory.} Once per session, become a living archive for one scene: gain +3 dice to historical knowledge, and you may ask the GM one question about a forgotten event or debt. The answer is always true. After the scene, you gain a permanent **Debt Scar** (your eyes are flecked with grey, like granite) but you may also, once per arc, ignore the time limit on any obligation – you simply outwait it, and the creditor moves on to more impatient debtors.
\end{tcolorbox}