% ============================================================================
% PATRON ENTRY: THE NINTH --- BEYOND COMPREHENSION (INFINITE INFORMATION & UNKNOWABLE TRUTHS)
% ============================================================================

\subsection{The Ninth --- Beyond Comprehension (Infinite Information \& Unknowable Truths)}

\subsubsection*{Lore}
Beyond the eighth figure of Sacred Geometry lies a rim no compass will hold. The Ninth is not absence but surplus: knowledge poured past the lip of any vessel, truth that overgrows the frame that seeks to bind it. Those who bend the knee at its lectern become conduits for impossible reckonings --- glimpsing patterns that cannot be kept whole inside a single skull. Each revelation purchases power at the price of living comfortably within the world's common measures.

The Ninth does not explain, define, or warn. It \textit{overflows}. To invoke it is to step past the eighth threshold --- the one you did not know was there --- and find that on the other side, the cup forgets its shape, the bell rings a note that was never cast, and the answer to your question arrives as a question you had not yet learned to ask.

\begin{quote}
``Eight figures bind the world; the Ninth overflows it. Drink, and remember that the cup forgets its shape before the song is done.''
\end{quote}

\subsubsection*{Domain Focus}
\begin{itemize}
  \item \textbf{Hyperdimensional Knowledge:} reckonings that do not fit a single plane or proof
  \item \textbf{Incomprehensible Truths:} insights that demand the seer be changed to hold them
  \item \textbf{Information Overflow:} boon and peril of minds run past their brim
  \item \textbf{Conceptual Transcendence:} thinking beyond the usual ledgers and names
  \item \textbf{Threshold Breaking:} stepping past the eighth boundary into what should not exist
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}

A Runekeeper sworn to the Ninth does not bind a living familiar. Their \textbf{Thiasos} is a \textit{pattern given will}: a sentient compass that points to contradictions, a brass abacus whose beads click in impossible sequences, or a mirror that reflects questions instead of faces. This entity can be housed in a small object (a lens, a pendulum, a blank book). It requires no food, but it demands new information: each week the Runekeeper must feed it a secret, a riddle, or a half-finished proof. If neglected for a month, the Thiasos begins to whisper false answers.

The \textbf{Codex} of the Ninth is never a simple grimoire. It may be a slate of fractally etched stone, a bundle of knotted threads, or the inside of the Runekeeper's own eyelids. Because the Ninth's truths resist being fixed in one place, a Codex requires upkeep: the Runekeeper must spend an hour each week erasing and rewriting a single page, or the Rites stored there become garbled (treat as Compromised until restored).

\subsubsection*{Patron's Gift (Imbuement)}

Once per scene as an action, a Runekeeper of the Ninth may touch a held item (weapon, tool, or writing implement) and speak a paradox. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Lore} or \textit{Investigation} rolls, and its first use each scene ignores one Story Beat generated by a 1 (the item's impossible nature briefly confuses the Weave). After the scene, the user marks \textbf{+1 Obligation} to the Ninth.

\subsubsection*{Rites of the Ninth}

\paragraph{Rite of the Ninth Figure} (Basic, 4 XP) \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[THRESHOLD]} \\
\textbf{Materials:} A chalk figure with an ``impossible'' joint, or a reckoning worked toward infinity on slate. \\
\textbf{Effect:} Glimpse past the ordinary frame of measure. Gain \textbf{+1 die} to \textit{Lore} or \textit{Investigation} when wrestling complex systems or many-layered problems. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Flooded Scriptorium} (Basic, 5 XP) \hfill \\
\emph{Scene; Near; Yes.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[SPEED]}, \texttt{[OVERFLOW]} \\
\textbf{Materials:} An overburdened ledger-hall, stacked codices, a bell that tolls the copying-hour without cease. \\
\textbf{Effect:} Read faster than hands can turn pages. Ask three questions upon any topic that leaves a written wake; gain \textbf{+2 dice} to rolls of analysis this scene. \\
\textbf{Push It:} The tide of knowing overruns its banks --- suffer \textbf{Fatigue 1} and mark \textbf{1 SB (Diamonds)} as overflow breeds complication. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Hypercognitive Lattice} (Advanced, 8 XP) \hfill \\
\emph{Scene; Self; Yes.} \\
\textbf{Tags:} \texttt{[BOOST]}, \texttt{[THRESHOLD]}, \texttt{[OVERFLOW]} \\
\textbf{Materials:} A brass-thought lattice, an orrery of inked threads, or a draught that sharpens memory. \\
\textbf{Effect:} Widen the mind's mouth. For this scene, treat your \textit{Wits} as \textbf{+1} higher for information-processing rolls; you may track several intricate streams at once.

\noindent\textbf{Memory Anchor:} The insight is overwhelming. You must write a brief summary of what you learned during downtime (a short scene or a full hour of transcription). Without this anchor, the bonus fades at the end of the scene and cannot be recovered.

\textbf{Push It:} The opening does not quite close: keep the boon, but suffer \textbf{--1 die} to single-task focus; mark \textbf{1 SB (Clubs)}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Breaking Glyph} (Advanced, 7 XP) \hfill \\
\emph{Extended; Self; Yes.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[THRESHOLD]} \\
\textbf{Materials:} A vigil in dim light, a water-timer, and a problem sworn not to be left half-seen. \\
\textbf{Effect:} Take one pattern whole. Choose a complex system, theory, or design and gain perfect internal grasp of it; receive \textbf{+3 dice} to related knowledge rolls for the \textit{next exchange} (one action). The insight is too vast to hold for longer. The GM gains \textbf{1 SB} as the knowledge strains the mind. \\
\textbf{Push It:} Your tongue cannot carry what your mind now holds: you cannot clearly teach or translate this insight; mark \textbf{1 SB (Hearts)}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Exchange.

\paragraph{Rite of the Endless Stacks} (Master, 13 XP) \hfill \\
\emph{Scene; Zone; Yes.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[TRUTH]}, \texttt{[OVERFLOW]}, \texttt{[THRESHOLD]} \\
\textbf{Materials:} A hall of mirrored shelves, wax-tablets without end, whispering leaves of vellum. \\
\textbf{Effect:} Raise a locus of overfull knowing. All present gain \textbf{+2 dice} to information-work; knowledge may be drawn from unlikely sources. Start a \textbf{[6] Information Overload} timer --- as it fills, participants strain and fray. \\
\textbf{Push It:} The Stacks wake and demand tithe in understanding or memory; mark \textbf{+2 Obligation}. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Ninth Revelation} (Master, 14 XP) \hfill \\
\emph{Extended; Self; Yes.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[THRESHOLD]}, \texttt{[TRANSFORM]}, \texttt{[FATE]} \\
\textbf{Materials:} A fast from easy speech, a circle of unbroken figures, and vows to see past comfort. \\
\textbf{Effect:} Glimpse a shard of Ninth truth. Choose one: perfect command of a knotty system for the next exchange; a single forecast of chaotic motion (one yes/no question answered by the GM); or comprehension of an ``impossible'' relation (the GM must answer one question truthfully). You must first succeed on an \textit{Investigation} or \textit{Lore} roll against a DV set by the GM (minimum 4). On a failure, the insight is garbled --- you gain no benefit, but mark \textbf{1 SB (Diamonds)} as the knowledge recoils. \\
\textbf{Push It:} The shard sets in you like glass: gain \textbf{Transcendent Understanding} (see Corruption) and mark \textbf{+2 Obligation}. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Exchange.

\subsubsection*{Cantors \& Cults of the Ninth}

\textbf{The Cantor's Song.} A Cantor of the Ninth sings in \textit{phrases that fold back on themselves} --- melodies that seem to have no beginning or end, harmonies that shift key in ways that should be impossible. Their voice is used to encrypt a message into a tune that only the intended listener can hear, to unravel the logic of a paradox, or to lull a library's guardian spirit into a momentary confusion. The Ninth's Cantors are often found in great archives, humming while they work.

\textbf{The Cult of the Unfinished Index.} The Ninth's cults meet in libraries, scriptoria, and the back rooms of printing houses --- but the innermost circle, the \textit{Unfinished Index}, has a stranger practice: each member carves a single extra joint into a finger of their writing hand (a ninth knuckle that does not bend). They believe that the extra joint holds a question they have not yet learned to ask. A ritual involves each member writing a fact they have never told anyone on a slip of paper, then burning it. The ashes are swept into a single book --- the \textit{Growing Ledger} --- whose pages multiply each year and \textit{whisper} at night, adding entries no one wrote. Their creed is stamped inside the cover: \textit{``Every answer births nine new questions. This is not a flaw. This is the shape of truth.''} Outsiders are welcome to ask a question, but they will leave with three they did not bring --- and sometimes, a faint, extra knuckle that fades by morning.

\subsubsection*{Witchcraft of the Ninth (Lay / Knowledge)}

A witch who walks with the Ninth is a \textit{glyph-carver} or \textit{pattern-weaver} --- one who crafts knowledge into physical form: a book that writes itself, a map that shows only the routes the reader does not yet know, or a code that cannot be broken because it was never written. Their magic works through the tools of information: a quill that writes in letters that shift meaning with the light, a seal that stamps ``APPROVED'' onto documents that were never submitted, or a lens that reveals hidden writing. Their craft is never about forging metal; it is about folding paper, spilling ink, and stepping past the eighth line of a proof into the white space where the answer is not written because the question was wrong.

\textbf{The Witch's Tool: The Paradox Quill.} A quill carved from the bone of a bird that never lived, dipped in ink that dries clear. Once per session, the witch may write a single sentence on a surface (a wall, a scroll, a forehead). The sentence is true --- but only for the next hour. After that, it becomes a lie, or fades, or was never there at all. The GM determines the exact effect, but the truth is always useful for that hour.

\textbf{A Hedge Gift: Glyph of Unfolding (Basic, 4 XP).} Once per scene, the witch may draw a single geometric symbol on a surface (a door, a lock, a piece of paper) while speaking a question aloud. The symbol ``unfolds'' into a diagram that answers the question --- but the diagram is always incomplete, cryptic, or reveals something the seeker did not wish to know. The witch gains \textbf{+1 die} to any roll that uses the diagram's information.

\subsubsection*{Corruption of the Ninth}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Hyperawareness:} +1 die to \textit{Notice} when patterns whisper that they matter. & \textbf{Starved for Text:} --1 die in barren halls with little record or sign; you itch for a stream to read. \\
2 & \textbf{Pattern Recognition:} Once/scene, see connections others miss within tangled accounts and signs. & \textbf{Scattered Threads:} Thoughts run in parallel; --1 die to sustained, single-line focus. \\
3 & \textbf{Data Synthesis:} Once/session, fuse vast reckonings at a glance into a usable answer. & \textbf{Tongue-Tied by Wonder:} --2 dice to teach or clearly explain your deepest insights. \\
4 & \textbf{Dimensional Insight:} Once/scene, take a problem from an ``impossible'' angle and treat a knot as if already loosened. & \textbf{Bent Sense:} Ordinary space and tasks feel ill-fit; --1 die to plain spatial or mundane labors. \\
5 & \textbf{Infinite Processing:} Once/session, render perfect analysis of any information system or record-trail. & \textbf{Concept Drift:} Common categories slip; --1 die to easy company and familiar talk. \\
6+ & \textbf{Ninth Awareness:} Once/session, look from outside the frame; for a heartbeat, the world aligns. & \textbf{Transcendent Fragmentation:} Mark +3 Obligation. You exist partially out of phase. Once per session, you may phase through a solid obstacle (wall, door, person) as a move action, but after doing so you cannot physically interact with objects for one exchange. The GM gains 1 SB as reality shivers. \\
\end{tblr}

\subsection*{Playstyle Notes}
The Ninth favors the patient reader of tangled ledgers and sky-signs alike. It rewards audacity in thought and the courage to let understanding change you. Its gifts turn riddles into roads --- but every road leads past a threshold you did not know existed. To serve the Ninth is to accept that you will never again fit comfortably inside the frame you once called home. Intrusions often take the form of sudden, inexplicable knowledge (a face you recognize from a dream, a word in a language you have never learned), or a question posed by a stranger who seems to know exactly what you were thinking. Refusing an intrusion does not bring punishment --- merely the slow forgetting of the insight you were about to receive, leaving you with the sense that you once knew something important, but can no longer recall what.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Hyperdimensional Knowledge:} insights that escape a single page
  \item \textbf{Cognitive Expansion:} widening the mind to hold more than comfort allows
  \item \textbf{Pattern Transcendence:} seeing threads that run beneath names and weights
  \item \textbf{Information Synthesis:} binding many scraps into one true line
  \item \textbf{Conceptual Breakthrough:} stepping past the frame rather than painting within it
  \item \textbf{Threshold Overflow:} the power of the ninth step, the extra word, the unasked question
\end{itemize}