% --- Patron: Oath of Flame & Light (Dawn & Vows) ---
\subsection{Oath of Flame \& Light (Dawn \& Vows)}

\textit{Lore.} The Oath of Flame \& Light is no patron of half‑measures. Its fire names, binds, and burns---demanding that those who swear within its radiance stand openly, speak truly, and pay the cost of keeping their word. At dawn altars, the sworn kindle sparks of consecrated fire; in battle, they blaze as torches that hold back the night. To follow this Oath is to live in public truth, with no shadow to hide in and no retreat from the vow once spoken.

The Oath appears in many guises across cultures, and is the patron most commonly called upon by the magic‑users of the Church of Flame and the Temple of Light -- as well as any fire‑ or light‑based faith or order:
\begin{itemize}
  \item \textbf{The Everflame} --- the unquenchable fire of promise, carried from shrine to shrine by wandering priests of the Church of Flame. Its censers never cool.
  \item \textbf{Adur}, the Light of Aeler --- worshiped as the sun's church‑deity, whose flame consecrates oaths and consumes falsehood. Dwarven templars strike their oaths into stone with a brand.
  \item \textbf{Adar}, the Vilikari Dawnfire --- invoked as a war‑god of vows and vengeance, whose rising light marks those who break faith. Cavalrymen light a single arrow before battle.
  \item \textbf{The Watchfire of Vhasia} --- a saint's flame that never gutters, tended by a silent order of healers who swear to harm no one -- and to let no one die alone.
  \item Others know the Oath only as \textit{the Flame} or \textit{the Lantern}, a nameless fire that still demands truth and punishes betrayal.
\end{itemize}

Wherever it manifests, the Oath of Flame \& Light remains the same unyielding power: to swear beneath its dawn is to bind one's self to a truth that cannot be hidden, forgotten, or undone.

\begin{quote}
``Swear in the light. Keep it, or the light will keep \emph{you}.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Sacred Vows:} Oath‑keeping, truth‑speaking, binding promises
  \item \textbf{Divine Justice:} Retribution, smiting evil, protecting the innocent
  \item \textbf{Radiant Power:} Healing light, purification, revealing truth
  \item \textbf{Dawn's Hope:} Renewal, protection, driving back darkness
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to the Oath of Flame \& Light does not bind a familiar of darkness or silence. Their \textbf{Thiasos} is a small creature of dawn and radiance: a fire‑coloured salamander that basks on hot coals, a small brass‑plated lizard whose scales warm when a vow is true, a golden‑eyed owl that hoots only at sunrise, or a tiny phoenix fledgling that glows faintly in darkness. This creature requires no food, but it must be present when the Runekeeper renews their daily vows (a short prayer at dawn). If neglected for a week, it becomes \textit{Compromised} (its glow flickers or goes out, imposing –1 die on all social rolls involving oaths or truth for its bearer until a new vow is sworn in its presence).

The \textbf{Codex} of the Oath is never a book of ink and paper. It may be a brass scroll etched with commandments that glow when read aloud, a set of consecrated candles whose flames spell out Rites, or a sun‑stone carved with runes that only appear when touched by direct sunlight. Because the Oath’s truths are written in light, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week in dawn light, re‑reading each Rite aloud, or the engravings fade (treat as Compromised until re‑consecrated at dawn).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of the Oath of Flame \& Light may touch a held item (weapon, shield, or badge) and whisper a vow of protection. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Defend} or \textit{Resist} rolls, and the first time each scene the bearer would be forced to break a sworn oath by magical compulsion, they may ignore it (the flame burns through the coercion). After the scene, the user marks \textbf{+1 Obligation} as the flame’s warmth lingers on their skin.

\subsubsection*{Rites of the Oath of Flame \& Light}

\paragraph{Rite of Kindle Vow (Low, 4 XP)} \hfill \\
\emph{Action; Self/Ally; Standard Push} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[RADIANT]} \\
\textbf{Materials:} A glass ampoule of consecrated flame. \\
\textbf{Effect:} Declare a short vow for this scene (e.g., "protect the villagers," "speak only truth"). The bearer gains \textbf{+1 die} to actions fulfilling it. \\
\textbf{Push It:} The first hesitation forces \textbf{1 SB (Hearts)} on the bearer. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of Lay on Hands (Low, 5 XP)} \hfill \\
\emph{Instant; Touch; Standard Push} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[PURIFY]}, \texttt{[RADIANT]} \\
\textbf{Materials:} A bare palm, a whispered vow. \\
\textbf{Effect:} Cleanse an affliction, downgrade Harm by 1, or remove 1 Fatigue. For curses or poisons, test Spirit+Resolve (DV set by fiction). \\
\textbf{Push It:} The target gains \textbf{+1 die} to their next Resist roll this scene; you mark Exposure +1. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of Sunlit Parley (Standard, 8 XP)} \hfill \\
\emph{Scene; Zone; Standard Push} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[COMMAND]}, \texttt{[RADIANT]} \\
\textbf{Materials:} A vow‑ring engraved with sunrise. \\
\textbf{Effect:} Establish terms in open light. Honest persuasion gains \textbf{+1 die}; deceit suffers \textbf{--1 die}. Create a 6‑segment \textit{Parley Accord} clock. \\
\textbf{Push It:} Demand one public answer; evasion forces \textbf{1 SB (Hearts)} on the evader. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of Radiant Smite (Standard, 9 XP)} \hfill \\
\emph{Action; Self; Standard Push} \\
\textbf{Tags:} \texttt{[STRIKE]}, \texttt{[BURN]}, \texttt{[RADIANT]} \\
\textbf{Materials:} A consecrated spark on a weapon or badge. \\
\textbf{Effect:} Your next melee strike flares with dawnfire. Upgrade Effect by one step, add +1 Burn Harm, or force \textbf{1 SB (Spades)}. Against undead, oath‑breakers, or outsiders: oath‑breakers suffer –1 die, outsiders gain +1 Exit Tally. \\
\textbf{Push It:} On a hit, the burst drives back enemies in Close (worse Position); mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of Purge the Shadow (Standard, 10 XP)} \hfill \\
\emph{Instant; Near; Standard Push} \\
\textbf{Tags:} \texttt{[DISPEL]}, \texttt{[REVEAL]}, \texttt{[RADIANT]} \\
\textbf{Materials:} A shattered consecrated spark. \\
\textbf{Effect:} Reveal illusions and suppress one ongoing glamour or curse in Near range. Creates a 4‑segment \textit{Purity's Light} clock. \\
\textbf{Push It:} Brand the source with a visible tell for this arc; mark \textbf{1 SB (Diamonds)}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of Covenant Blaze (High, 13 XP)} \hfill \\
\emph{Scene; Zone; High Push} \\
\textbf{Tags:} \texttt{[WARD]}, \texttt{[FORCE]}, \texttt{[RADIANT]}, \texttt{[COMMAND]} \\
\textbf{Materials:} A brazier lit while three names are spoken. \\
\textbf{Effect:} The sworn within are haloed: they gain \textbf{+1 die} to keep their oath; aggressors suffer \textbf{--1 die} if violating the terms. Oath‑breakers suffer \textbf{2 SB (Hearts/Spades)} and Harm 1 (Burn). \\
\textbf{Push It:} Sanctify a threshold with one beat of \texttt{[WARD]} against oath‑breakers; mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 7 segments.

\subsubsection*{Cantors \& Cults of the Oath of Flame \& Light}

\textbf{The Cantor's Song.} A Cantor of the Oath sings in \textit{clarion tones} -- melodies that sound like a trumpet at sunrise, like a bell calling the faithful to prayer. Their voice is used to swear in a witness, to consecrate a blade with a hymn, or to sing a funeral pyre into a beacon. The Oath's Cantors are often found in military camps and pilgrimage roads, their songs keeping oaths fresh and courage hot.

\textbf{The Cult of the Unquenched Flame.} The Oath's cults gather at dawn, in open courtyards or on hilltops facing east. The Unquenched Flame is a fellowship of paladins, healers, and simple folk who have sworn to keep one another honest. A ritual of the Cult involves each member speaking a vow aloud as the sun crests the horizon, then passing a lit candle to the next -- the flame must not be allowed to go out until all have spoken. Their creed is written in charcoal above every hearth: \textit{"The flame that reveals the truth also warms the hands that keep it."} Outsiders are welcome to swear a temporary vow and share the light -- but an oath sworn at dawn is binding until sunset.

\subsubsection*{Witchcraft of the Oath of Flame \& Light (Monastic / Scholarly)}

A witch who walks with the Oath is often a \textit{consecrated scribe}, a \textit{censer‑bearer}, or a \textit{vowed healer} – a nun, monk, or scholar of one of the great flame‑temples. Their magic is inseparable from liturgy, prayer, and the careful tending of sacred spaces. They illuminate manuscripts with gold leaf that glows in the presence of lies, brew unguents from ashes of the everflame, and copy oath‑books whose letters burn when a vow is broken. In Aeler, a True‑Mason might carve a “prayer keystone” into a basilica wall, so that every confession spoken there warms the stone. In Vhasia, a silent sister might stitch a healing shroud that records the wounds it has mended.

\textbf{The Witch’s Tool: Prayer Beads of the Unbroken Vow.} A string of nine beads carved from olive wood and fired in a basilica’s censer. Each bead represents a clause of a sacred oath. Once per session, the witch may slip the beads around a door handle, a prisoner’s wrist, or a contract scroll. While the beads remain, any creature that knowingly violates an oath sworn under that threshold suffers a mild burn (1 Fatigue) as the beads glow faintly. The beads may also be used to count a silent vow; the witch may reroll one failed Resolve roll when upholding that vow.

\textbf{A Hedge Gift: Sealed Benediction (2 XP).} Once per scene, the witch may lay a hand on a willing creature’s forehead and speak a short, formal blessing (e.g., “May the flame keep your word”). The target gains \textbf{+1 die} on their next roll to resist fear, despair, or magical coercion. The blessing lasts until the next dawn. The witch must write the blessing on a scrap of parchment and burn it; the smoke carries the prayer.

\subsubsection*{Oath of Flame \& Light Corruption Manifestations}

\begin{tblr}{
  colspec = {Q[c,1.8cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Level & Benefit & Cost / Quirk \\
1 & \textbf{Oathbound Strength:} +1 die when upholding vows or defending innocents. & \textbf{Rigid Honour:} Must uphold vows even when disadvantageous; --1 die to flexible actions. \\
2 & \textbf{Radiant Sight:} Once/scene, +2 dice to detect lies or corruption. & \textbf{Blinding Truth:} --1 die to subtlety or deception attempts. \\
3 & \textbf{Holy Flame:} +1 die to melee against undead, outsiders, or oath‑breakers. & \textbf{Burden of Light:} 1 Fatigue when concealing identity or working in darkness. \\
4 & \textbf{Unwavering Resolve:} Once/session, treat a failed Resolve/Command as success (mark 1 SB). & \textbf{Absolutist Stance:} --1 die in morally ambiguous situations. \\
5 & \textbf{Dawn's Benediction:} Once/session, heal all allies in Near of 1 Fatigue and minor Conditions. & \textbf{Beacon's Call:} Your aura reveals you; enemies seeking you gain +1 die. \\
6+ & \textbf{Avatar of the Oath:} Once/session, become a living covenant (+2 dice to protection/justice actions). & \textbf{Radiance's Price:} Mark +2 Obligation; breaking any vow inflicts Harm 2 (Burn). \\
\end{tblr}

\subsection*{Playstyle Notes}
The Oath of Flame \& Light excels as a “paladin” patron, rewarding vow‑keeping and truth‑speaking with potent defensive and smiting abilities. Followers become living embodiments of their oaths, shining beacons against deception and darkness. The corruption progression naturally leads toward becoming an uncompromising force for justice, though potentially at the cost of flexibility and subtlety. Ideal for characters who embody conviction, protection, and radiant power.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Sacred Vows:} oath‑keeping, truth‑speaking, binding promises
  \item \textbf{Divine Justice:} retribution, smiting evil, protecting the innocent
  \item \textbf{Radiant Power:} healing light, purification, revealing truth
  \item \textbf{Dawn's Hope:} renewal, protection, driving back darkness
\end{itemize}