% --- Patron: Oath of Mercy and Grace (Healing & Forgiveness) ---
\subsection{Oath of Mercy and Grace (Healing \& Forgiveness)}

\textit{Lore.} The Oath of Mercy and Grace has no founder, no scripture, no single temple. It began in a hundred places—a battlefield where a medic bound wounds without asking whose side; a plague village where a stranger stayed to dig graves; a debtor’s cell where a guard looked away and said, “I saw nothing.” Those who swear the Oath do not worship a power; they embody a principle: that every act of care creates an obligation—not to the giver, but to the next person in need. You do not pay back. You pay forward. The mercy spirals.

But mercy is not weakness. The spiral does not forget. Those who strike a healer, who violate a sanctuary, who refuse the offered cup—they find that grace has teeth, and forgiveness carries a price.

Among the peoples of Fate's Edge, the Oath of Mercy and Grace is the patron of healers, hospice workers, and those who shelter the helpless. In Aelaerem, the Oath is kept by the \textit{Grey Spiral}, a fellowship of lay healers who ask no payment. In Viterra, the \textit{Hospitaller Knights} swear to defend the sick without drawing swords. In Zakov, the \textit{Tide‑Watchers} pull drowning sailors from the sea and ask only that the rescued help one stranger in turn. Wherever mercy is given without expectation of return, the spiral turns.

\begin{quote}
``You ask who owes what. We ask who needs what. The ledger is not the same as the meal. But woe to the one who mistakes the server for a servant.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Restorative Work:} healing body, mind, and spirit; mending what is broken.
  \item \textbf{Spiral Mercy:} paying obligation forward through acts of care.
  \item \textbf{Sanctuary:} sheltering the hunted, feeding the hungry, accepting no payment—but demanding respect for the threshold.
  \item \textbf{Unburdening:} releasing others (and oneself) from the weight of unpaid debts, and binding aggressors to that same weight.
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to the Oath of Mercy and Grace does not bind a familiar of claw or fang. Their \textbf{Thiasos} is a small creature of gentle presence and quiet vigilance: a white dove that perches on the shoulder and coos when a lie is spoken in a place of sanctuary, a grey cat that curls on the laps of the dying and purrs away their fear, a small hedgehog that rolls into a ball to protect the vulnerable, or a firefly that glows warmly when a vow of mercy is kept. This creature requires no food, but it must witness an act of care (feeding the hungry, binding a wound, sheltering a fugitive) each day. If neglected for a week, it becomes \textit{Compromised} (it hides or refuses to approach the sick, imposing –1 die on all healing and social rolls involving mercy for its bearer until an act of care is performed in its presence).

The \textbf{Codex} of Mercy and Grace is never a book of spells alone. It may be a grey wool cord braided with nine knots, each knot tied while speaking a name of someone who was healed; a clay bowl fired from the clay of a hospice hearth, its inner surface inscribed with vows; or a ledger of debts forgiven, each page sealed with a thumbprint of ash. Because the spiral is written in acts, not ink, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week untying and retying each knot (or re‑firing the bowl, or reviewing a page of forgiven debts), or the Rites become uncertain (treat as Compromised until restored by an act of mercy).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of the Oath of Mercy and Grace may touch a held item (weapon, shield, or bandage) and whisper a word of unburdening. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Medicine} or \textit{Comfort} rolls, and the first time each scene the bearer would take Harm while defending a helpless creature, they may reduce that Harm by 1 (to a minimum of 0) as the spiral catches the blow. After the scene, the user marks \textbf{+1 Obligation} as the weight of another’s pain briefly rests on them.

\subsubsection*{Rites of the Oath of Mercy and Grace}

\paragraph{Rite of the Shared Bowl (Low, 4 XP)} \hfill \\
\emph{Action; Touch; Standard Push} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[BOND]}, \texttt{[RESTORATION]} \\
\textbf{Materials:} A clay bowl of clean water, shared with a willing creature. \\
\textbf{Effect:} You and the creature each clear 1 Fatigue. The creature gains +1 die on their next roll to resist fear or despair. \\
\textbf{Push It:} Extend the benefit to all creatures who drink from the bowl (up to six). Mark +1 Obligation. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of Healing Waters (Low, 5 XP)} \hfill \\
\emph{Instant; Touch; Standard Push} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[PURIFY]}, \texttt{[RESTORATION]} \\
\textbf{Materials:} Water from a blessed source (a spring, a well, or a river where the Oath has been invoked), poured onto a wound. \\
\textbf{Effect:} Remove 1 Harm from a touched creature. If the wound is cursed or poisoned, the GM may call for a Spirit+Lore test (DV 4). \\
\textbf{Push It:} Also remove one Condition (e.g., Poisoned, Cursed, Blinded). Mark +1 Obligation. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Vow of Pacifism (Low, 4 XP)} \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[PROTECT]}, \texttt{[DEBT]}, \texttt{[FEAR]} \\
\textbf{Materials:} A white cord tied around the weapon hand. \\
\textbf{Effect:} You swear not to kill for the scene. In exchange, you gain \textbf{+1 Armor} and \textbf{+1 die to all resist rolls}. If a creature harms you while this vow is active, they immediately gain a \textbf{Promise Timer [4]} named “Mercy’s Reckoning.” When that timer fills (by them continuing violence), they suffer Harm 1 (Stress) as their own aggression turns inward. \\
\textbf{Push It:} Extend the protective effect to one ally within Near. Mark +1 Obligation. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Unburdened Circle (Standard, 8 XP)} \hfill \\
\emph{Scene; Zone; Standard Push} \\
\textbf{Tags:} \texttt{[DEBT]}, \texttt{[BOND]}, \texttt{[RESTORATION]}, \texttt{[AREA]} \\
\textbf{Materials:} At least three willing participants, a circle drawn in ash or flour, a single candle. \\
\textbf{Effect:} Each participant clears 1 Obligation (to any patron or debt). One participant of the GM’s choice gains a permanent String: \textit{Witnessed by the Circle} – they may call on any other participant for one act of care, once per campaign. \\
\textbf{Push It:} The GM gains 3 Story Beats (the universe notes that a community of mutual aid is forming). Mark +1 Obligation. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Sheltered Hearth (Standard, 9 XP)} \hfill \\
\emph{Scene; Zone; Standard Push} \\
\textbf{Tags:} \texttt{[WARD]}, \texttt{[PROTECT]}, \texttt{[DEBT]}, \texttt{[AREA]}, \texttt{[FEAR]} \\
\textbf{Materials:} A door or threshold, a lit candle, a whispered vow of sanctuary. \\
\textbf{Effect:} For one night, any fugitive, debtor, or refugee who crosses that threshold cannot be pursued or harmed within the zone. The sanctuary holds until dawn. Any creature that violates this sanctuary (by attacking or forcibly entering) gains a \textbf{Promise Timer [6]} named “Sanctuary Breach.” When full, they suffer a Curse chosen by the GM (e.g., “doors close in your face,” “your name is forgotten by hosts”). \\
\textbf{Push It:} Extend the sanctuary to a full day. Mark +1 Obligation. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Shield of the Unburdened (Standard, 8 XP)} \hfill \\
\emph{Instant; Near; No.} \\
\textbf{Tags:} \texttt{[PROTECT]}, \texttt{[DEBT]}, \texttt{[TRANSFER]}, \texttt{[FEAR]} \\
\textbf{Materials:} A grey cloak or a bandage wrapped around your own arm. \\
\textbf{Effect:} When an ally within Near would take Harm, you may interpose yourself and take that Harm instead. After doing so, the attacker must make a \textbf{Spirit+Resolve} test (DV 4) or gain a \textbf{Promise Timer [4]} named “Debt of Blood.” The timer ticks each time they refuse a genuine plea for mercy or aid. When full, they suffer Fatigue 2 as guilt manifests. \\
\textbf{Push It:} You may also redirect a Condition (e.g., Poisoned, Cursed) from the ally to yourself. Mark +1 Obligation. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Spiral's Witness (Standard, 10 XP)} \hfill \\
\emph{Instant; Near; Standard Push} \\
\textbf{Tags:} \texttt{[DEBT]}, \texttt{[RESTORATION]}, \texttt{[FEAR]}, \texttt{[TRUTH]} \\
\textbf{Materials:} A grey wool cord tied around the caster’s wrist. \\
\textbf{Effect:} Name a person crushed by debt or guilt. For one hour, they ignore all penalties from unpaid obligations (social, legal, or spiritual). They are not free; they are unburdened. If that person then commits an act of violence against an innocent, the unburdening reverses and they suffer double penalties. \\
\textbf{Push It:} The effect lasts until the next dawn. Mark +1 Obligation. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Covenant of Shared Suffering (High, 13 XP)} \hfill \\
\emph{Scene; Zone; High Push} \\
\textbf{Tags:} \texttt{[BOND]}, \texttt{[TRANSFER]}, \texttt{[HEAL]}, \texttt{[AREA]}, \texttt{[DEBT]} \\
\textbf{Materials:} A circle of nine grey stones, a bowl of water, a candle lit at each cardinal point, and a drop of blood from each participant. \\
\textbf{Effect:} All within the circle who accept the Oath’s blessing are linked. For the scene:
\begin{itemize}
  \item Damage taken by any linked ally may be split among any other linked allies (each chooses to accept 1 Fatigue or 1 Harm).
  \item When you heal a linked ally, you may also heal yourself for half the amount (round down).
  \item The first time a linked ally would be reduced to 0 Harm, you may sacrifice 1 Fatigue to keep them standing.
\end{itemize}
\textbf{Push It:} The covenant persists for a full day. Mark +2 Obligation. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Endless Spiral (High, 14 XP)} \hfill \\
\emph{Extended; Zone; High Push} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[RESTORATION]}, \texttt{[DEBT]}, \texttt{[AREA]}, \texttt{[FEAR]} \\
\textbf{Materials:} A circle of nine grey stones, a bowl of tears, and a candle lit at each cardinal point. \\
\textbf{Effect:} All within the circle who accept the Oath’s blessing clear 2 Fatigue and remove one Condition caused by despair or guilt. For the rest of the scene, any act of care performed within the zone costs the giver –1 Obligation (minimum 0). Additionally, any creature that harms a caregiver within the zone gains a \textbf{Promise Timer [6]} named “Spiral’s Debt.” When full, they suffer Harm 2 (Stress) as the weight of their cruelty crushes them. \\
\textbf{Push It:} The blessing persists for a full day. Mark +2 Obligation. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 8 segments.

\subsubsection*{The Spiral Strikes Back (8 XP, Major)}
\emph{Prerequisites: Oath of Mercy and Grace, Spirit 3+.} \\
You have learned that mercy can be a weapon. When a creature with an active Promise Timer (from any of your Rites) attacks you or an ally under your protection, you may spend 1 Boon to immediately advance that timer by 2 segments. If this fills the timer, the creature suffers the consequence immediately (Harm, Condition, or Curse as described). You may only use this once per scene.

\subsubsection*{Corruption Perk (Tier 4+): Mercy's Edge}
Once per session, when you would reduce a sentient enemy to 0 Harm, you may instead spare their life and bind them with a \textbf{Promise Timer [6]} named “Spared.” While this timer is active, they cannot knowingly harm you or those you have explicitly named as under your protection. They may work off the timer by performing three acts of genuine care for others (GM discretion). If they break the bond, the timer fills instantly and they suffer Harm 2 (Stress) – the spiral closes.

% --- Hospitaller Knight of the Grey Spiral (tcolorbox) ---
\begin{tcolorbox}[colback=black!5,colframe=blue!50!black,title=\textbf{Hospitaller Knight of the Grey Spiral}]
\emph{Mercy without steel is a prayer. Steel without mercy is a curse. We carry both – but we sheath the blade before we draw it.}

\vspace{2mm}
\noindent\textbf{Prerequisites:} Spirit 3+, Body 2+, Rite of the Sheltered Hearth or Rite of the Shield of the Unburdened.

\vspace{2mm}
\noindent\textbf{Hospitaller Techniques (choose as Talents):}

\begin{tblr}{
  colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.3,l]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Technique & Effect & Cost & Req. \\
\textbf{Grey Ward} & When you take the \textit{Defend} action, you also grant an adjacent ally +1 Armour until your next turn. & Passive & Body 3, Melee 2 \\
\textbf{Shield of the Unburdened} & Once per scene, as a reaction, redirect a melee attack aimed at an adjacent ally to yourself. The attack automatically hits you. & 1 Fatigue & Spirit 3 \\
\textbf{Mercy's Pace} & You ignore the first point of Fatigue each scene caused by exhaustion (forced marches, lack of sleep, extreme weather). & Passive & Spirit 3, Tier II \\
\textbf{Spiral Counter} & When a creature with an active Promise Timer attacks you, you may spend 1 Boon to advance that timer by 1 segment immediately, before the attack resolves. & 1 Boon & Tier III \\
\textbf{Hospitaller's Vow} & You may use the \textit{HEAL} cantrip on any ally within Near range, not just touch. & +1 Obligation per use & Spirit 3, HEAL cantrip \\
\textbf{Unbroken Cord} & When you use a Promise Timer from any of your Oath rites, it can only be advanced by the GM if you fail a Spirit+Resolve test (DV 4). Otherwise, the timer ticks only when you or an ally is harmed. & Passive & Tier III, Spirit 4 \\
\end{tblr}

\vspace{2mm}
\noindent\textbf{Master Technique (Epic Talent, 12 XP):}
\textit{The Spiral’s Final Witness.} Once per session, when a creature with an active Promise Timer (from any of your Oath rites) would deal a killing blow to an ally, you may interpose yourself. The blow is redirected to you, and you suffer the Harm instead. After the attack, the Promise Timer fills immediately, and the attacker suffers the consequence (Harm, Condition, or Curse). You cannot use this talent again until you have spent a downtime scene in meditation or prayer.
\end{tcolorbox}

\subsubsection*{Cantors \& Cults of the Oath of Mercy and Grace}

\textbf{The Cantor’s Song.} A Cantor of the Oath sings in \textit{harmonies of the unburdened} – melodies that lift without rising, harmonies that wrap around a wound like a clean bandage. Their voice is used to calm the panicked, to ease the pain of the dying, or to sing a debtor gently into the spiral. Mercy’s Cantors are often found in hospices, refugee camps, and the halls of the condemned, their songs a reminder that no debt is too heavy to be carried by more than one pair of hands.

\textbf{The Cult of the Grey Spiral.} The Oath’s cults are not secret – they are quiet. They gather in hospices, soup kitchens, and the prayer rooms of hospitals. The Grey Spiral is a fellowship of healers, hospice workers, and those who have learned that mercy is a muscle that must be exercised. A ritual of the Spiral involves each member tying a grey wool cord around their wrist (or a patient’s wrist) while speaking a single act of care they received from a stranger. The cord is then cut and burned, and the ashes are scattered in a place where the hungry gather. Their creed is embroidered on every grey cloak: \textit{"You do not pay back. You pay forward. The spiral is patient."} Outsiders are welcome to sit in silence, but they must accept a cup of water and speak a truth about a wound they carry.

\textbf{The Hospitaller Knights of the Grey Spiral.} A militant order within the Oath who believe that mercy must be defended. They carry no swords—only shields, staves, and the grey wool cord. They are sworn to protect healers, refugees, and the sick, and they have a fearsome reputation: \textit{“Strike the healer, and the spiral will remember. The debt will be collected, if not by us, then by the weight of your own cruelty.”}

\subsubsection*{Witchcraft of the Oath of Mercy and Grace (Lay / Monastic)}

A witch who walks with the Oath is a \textit{lay healer}, a \textit{hospice sister}, or a \textit{grey brother} – a nun, monk, or lay religious who tends the sick, comforts the dying, and offers sanctuary to the hunted. Their magic is not the flashy alchemy of a potion‑maker; it is the quiet, patient work of a wound that closes a day early, a fever that breaks at midnight, or a debtor who finds the courage to confess. Their tools are not hammers or chisels; they use grey wool cords, clay bowls, and the small brass bells that ring only when a prayer is answered.

\textbf{The Witch’s Tool: The Cord of the Unbroken Spiral.} A grey wool cord braided with nine knots, each knot tied in silence while speaking a name of someone who was healed. Once per session, the witch may touch the cord to a wounded creature (including themselves) and untie a single knot. The creature clears 1 Fatigue and, if they are suffering from a minor Condition (Fear, Shaken, Guilty), may test Spirit+Resolve (DV 3) to remove it. The knot cannot be retied until the next dawn. The cord can also be used to mark a threshold: tying it around a door handle or bedpost grants +1 die to all rolls made to protect the inhabitant for the night.

\textbf{A Hedge Gift: Benediction of the Spiral (2 XP).} Once per scene, the witch may lay a hand on a willing creature and whisper a short prayer (or a word of forgiveness). For the next hour, the creature feels lighter – they ignore the first point of Fatigue from exhaustion, and any social roll made to ask for mercy or forgiveness gains +1 die. The blessing ends if the creature commits a violent act against an innocent.

\subsubsection*{Oath of Mercy and Grace Corruption Manifestations}

\begin{tblr}{
  colspec = {Q[c,1.8cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Level & Benefit & Cost / Quirk \\
1 & \textbf{Soothing Presence:} +1 die to calm or comfort others. & \textbf{Burden of Care:} –1 die when ignoring suffering; must help or mark Fatigue. \\
2 & \textbf{Quick Bandage:} Once/scene, heal 1 Fatigue as a reaction. & \textbf{Empathic Bleed:} You feel the pain of those you heal; 1 Fatigue if you heal a major wound. \\
3 & \textbf{Sanctuary Sense:} +2 dice to detect sanctuary violations. & \textbf{Threshold Weary:} –1 die in places with no sacred boundary. \\
4 & \textbf{Spiral Witness:} Once/session, force an aggressor to test Resolve or be shaken. & \textbf{Promised Weight:} Must keep all promises to the vulnerable; breaking one costs 1 SB (Hearts). \\
5 & \textbf{Merciful End:} Once/session, grant a peaceful death to a willing creature (no roll). & \textbf{Echo of Loss:} Suffer 1 Fatigue when you cannot prevent a death. \\
6+ & \textbf{Spiral Incarnate:} Once/session, become a living knot; all Promise Timers in the scene advance toward fulfillment. & \textbf{Debt of Mercy:} Mark +2 Obligation; risk becoming unable to refuse a plea for help. \\
\end{tblr}

\subsection*{Playstyle Notes}
The Oath of Mercy and Grace supports both pure healers and \textbf{Hospitaller defenders}—characters who wade into battle not to kill, but to shield, to bind, and to turn violence back on itself as obligation. Your “teeth” are not fangs but \textit{leashes}: every enemy who harms you or your charges accrues a debt that will eventually cripple them. This works best in campaigns where the GM tracks Promise Timers and where enemies have names and faces—the spiral is less effective against mindless beasts, but against human foes, it is terrifying.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Restorative Work:} healing body, mind, and spirit
  \item \textbf{Spiral Mercy:} paying obligation forward through acts of care
  \item \textbf{Sanctuary:} sheltering the hunted, feeding the hungry, demanding respect for the threshold
  \item \textbf{Unburdening:} releasing others (and oneself) from the weight of unpaid debts
\end{itemize}

\subsection*{Rivalries}
\begin{itemize}
  \item \textbf{The Inquisitor Prime:} Mercy without judgment is heresy to the Prime. The two orders clash constantly over the fate of prisoners, heretics, and the “unforgivable.”
  \item \textbf{Malachai (Cruel Messenger):} Malachai offers cursed gifts; the Oath offers unburdening. Their followers are natural enemies—one feeds on despair, the other starves it.
\end{itemize}

% --- Monastic Tradition: Pacifist Monk (Oath of Mercy and Grace) ---
\begin{tcolorbox}[colback=black!5,colframe=green!70!black,title=\textbf{Pacifist Monk (Oath of Mercy and Grace)}]
  \emph{You do not pay back. You pay forward. The spiral is patient.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Spirit 3+, Presence 2+, Rite of the Shared Bowl or Cantor of Mercy and Grace.
  
  \vspace{2mm}
  \noindent\textbf{Debt–Resistant Frame.} Mercy is not a loan – it is a gift. Pacifist monks gain \textbf{+1 die} to resist any magical debt or obligation that would compel them to harm another or break an oath of non‑violence. If they succeed, the creditor’s demand is gently refused; the universe notes a kindness, and the Hollow’s attention briefly wanders.
  
  \vspace{2mm}
  \noindent\textbf{Techniques (choose as Talents):}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.4,l]},
    rowhead = 1,
    row{odd} = {bg=gray!10},
    row{1} = {bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Redirect Blow} & When an ally within Near would take Harm, you may interpose yourself and take that Harm instead. & 1 Fatigue & Body 3 \\
  \textbf{Healing Touch} & Remove 1 Fatigue or 1 Harm (not Condition) from a touched creature. Once per scene per target. & 1 Boon & Spirit 3 \\
  \textbf{Vow of Pacifism} & Swear not to kill for the scene. Gain +1 Armour and +1 die to all resist rolls. If a creature harms you while this vow is active, they gain a Promise Timer [4] named “Mercy’s Reckoning.” & 1 Fatigue & Presence 2 \\
  \textbf{Sheltered Hearth} & For one night, any fugitive or refugee who crosses a threshold you guard cannot be pursued or harmed within that zone. Sanctuary holds until dawn. & 1 Boon & Spirit 2 \\
  \textbf{Spiral Counter} & When a creature with an active Promise Timer attacks you, you may spend 1 Boon to advance that timer by 1 segment immediately, before the attack resolves. & 1 Boon & Tier II \\
  \textbf{Unburdened Circle} & Each participant clears 1 Obligation (to any patron or debt). One participant of the GM’s choice gains a permanent String: Witnessed by the Circle – they may call on any other participant for one act of care, once per campaign. & 1 Fatigue & Tier III \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Spiral Incarnate.} Once per session, become a living knot of mercy for one scene: all Promise Timers in the scene advance toward fulfillment, and any creature that harms an innocent under your protection immediately suffers Harm 2 (Stress). After the scene, you gain a permanent \textbf{Debt Scar} (a grey thread tied around your wrist – it never loosens, but it also never tightens). You may ignore one debt or obligation that would force you to abandon a plea for help – the creditor simply finds you already tending to someone else.
\end{tcolorbox}