% ============================================================================
% PATRON ENTRY: RAÉYN, MISTRESS OF THE SEA (TIDES & CHANGE)
% ============================================================================

\subsection{Raéyn, Mistress of the Sea (Tides \& Change)}

\subsubsection*{Lore}
Raéyn is the tempestuous goddess of the sea, the restless tide that carries news between shores and the promise of change between lives. She is mother to all who sail, her voice the wind that fills sails and her moods the storms that test every mariner's resolve. In Kahfagia, she is called the Storm-Queen; in Linn, the Tide-Mother; in the Brass Coast, the Salt-Widow. Her cult is syncretic, blending the fierce protectiveness of a mother goddess with the practical cunning of a harbor-master.

But Raéyn's heart is torn by her greatest tragedy: her son Khemesh, the Kraken of the Depths, who embodies the crushing inevitability of the ocean's dark heart. Where Raéyn brings change and opportunity, Khemesh brings the final, inescapable pressure that grinds all things to nothing. Sailors pray to Raéyn for safe passage and favorable winds, but whisper Khemesh's name when seeking to lay the dead to rest beneath the waves.

Raéyn is passionate, mercurial, and fiercely protective of those who respect her domain. She favors those who read currents, bargain with weather, and carry news between shores. But cross her, and the sea itself becomes your enemy: fair weather turns to fury, and every wave a judgment.

\begin{quote}
``Mark the tide, name your course, and trust the wave-road. But speak ill of Khemesh, and even I may let the deep take you.''
\end{quote}

\subsubsection*{Domain Focus}
\begin{itemize}
  \item \textbf{Tides and Currents:} change, opportunity, the rhythm of the sea
  \item \textbf{Maritime Trade and Travel:} safe passage, navigation, weather
  \item \textbf{Storm and Calm:} the duality of protection and wrath
  \item \textbf{Motherhood and Loss:} fierce love, grief for Khemesh, the weight of a divided heart
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}

A Runekeeper sworn to Raéyn does not bind a creature of dry land---they bind a companion of the tides. Their \textbf{Thiasos} is a small sea-creature or mythological being of the waves: a storm-petrel that appears before gales and can sense pressure drops, a gray seal pup that sings in harmonic tones audible only underwater, a hermit crab whose shell is etched with wave-runes, a small octopus that changes color to reflect the weather, or a sea-horse that can hover in still air for a heartbeat. This creature requires no food in the usual sense, but it must be immersed in salt water (or exposed to sea spray) each day. If kept dry for more than a day, it becomes \textit{Compromised} (it grows listless and its color fades; its bearer loses the benefit of \textit{Sailor's Instinct} until it is returned to the sea).

The \textbf{Codex} of Raéyn is never a book of dry paper. It may be a set of tide-charts etched on whalebone, a bundle of knotted cords that record the phases of the moon and the turn of currents, a conch shell whose interior spiral is inscribed with navigational verse, or a driftwood plank carved with wave-runes that only become visible when wet. Because the sea resists being fixed in writing, the Codex requires upkeep: the Runekeeper must spend an hour each week soaking it in salt water or tracing its symbols by the light of a tide pool, or the Rites stored there become faint (treat as Compromised until re-consecrated at the shore).

\subsubsection*{Patron's Gift (Imbuement)}

Once per scene as an action, a Runekeeper of Raéyn may touch a held item (weapon, tool, or piece of gear) and whisper a promise of safe passage. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Navigation} or \textit{Athletics} rolls when aboard a vessel or in water, and the first time each scene the bearer would suffer an environmental penalty from rough seas, wind, or rain, they may ignore it. After the scene, the user marks \textbf{+1 Obligation} to Raéyn.

\subsubsection*{Rites of Raéyn}

\paragraph{Rite of the Wave-Tide} (Basic, 5 XP) \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[MOVEMENT]}, \texttt{[NAVIGATION]}, \texttt{[BOOST]}, \texttt{[RISK]} \\
\textbf{Materials:} A knotted length of salt-twine brushed with seawater, or a shell held to the ear while facing the wind. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Tidemark's Blessing:} Treat slick, swaying, or water-slicked footing as stable for this scene. Gain \textbf{+1 die} on boarding, balance, or shipboard movement. Create a 4-segment \textit{Tide's Favor} timer to ignore one level of difficult terrain.
  \item \textbf{Whispering Currents:} Learn the safest near-term route across water or coastline, or gain \textbf{+1 die} to navigation checks for this scene. If Khemesh's influence is present, suffer --1 die from conflicting currents.
\end{itemize}
\textbf{Push It:} (Tidemark) Extend to one ally in Close for one beat, but generate \textbf{1 SB (Spades)} as the deck shifts. (Currents) Also learn the fastest route, but mark Exposure +1 (leaving a telltale wake). Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Changing Tide} (Advanced, 8 XP) \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[TERRAIN]}, \texttt{[AREA]}, \texttt{[MOVEMENT]}, \texttt{[PRESSURE]} \\
\textbf{Materials:} A handful of pebbles cast in a crescent. \\
\textbf{Effect:} Bias currents and water levels in the zone (must be water-adjacent). Those moving with the tide gain \textbf{+1 die}; those moving against suffer \textbf{--1 die}. Small craft must test to hold position. Create a 6-segment \textit{Tidal Influence} timer. \\
\textbf{Push It:} Brief surge or drawdown (one beat): open a ford or swamp a skiff. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Wave-Road Consecration} (Advanced, 7 XP) \hfill \\
\emph{Scene; Route; No.} \\
\textbf{Tags:} \texttt{[GUIDANCE]}, \texttt{[BOOST]}, \texttt{[BOND]}, \texttt{[FATE]} \\
\textbf{Materials:} Two sea-glass markers dropped overboard at start and end. \\
\textbf{Effect:} Consecrate a wave-road between two visible points. Allies gain \textbf{+2 dice} on travel, evade, or carry actions at sea. Designated pursuers suffer \textbf{--1 die} to intercept. One active wave-road at a time. Create an 8-segment \textit{Blessed Passage} timer. \\
\textbf{Push It:} Extend the route's favor to an adjacent leg for one beat; mark \textbf{+1 Obligation}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Storm-Queen's Hand} (Master, 13 XP) \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[WEATHER]}, \texttt{[AREA]}, \texttt{[COMMAND]}, \texttt{[FEAR]}, \texttt{[BOOST]} \\
\textbf{Materials:} A vial of rainwater gathered at three crossings. \\
\textbf{Effect:} Shape a storm-band over the zone (sea, shore, or open sky). Choose two modes at cast; switch one once per scene:
\begin{itemize}
  \item \textbf{Propulsion:} Vessel gains +1 band of movement per beat (or +1 Effect to maneuvers).
  \item \textbf{Concealment:} Veil of rain/spray; ranged targeting impaired; --1 die to hostile sighting.
  \item \textbf{Smite:} Once per beat, lash with wave or lightning as \texttt{[AREA]} hazard.
\end{itemize}
\textbf{Push It:} Add a third mode for one beat, then the GM spends 1 SB on collateral. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Mother's Wrath} (Master, 14 XP) \hfill \\
\emph{Extended; Zone; No.} \\
\textbf{Tags:} \texttt{[CURSE]}, \texttt{[DEBT]}, \texttt{[FEAR]}, \texttt{[AREA]}, \texttt{[FATE]} \\
\textbf{Materials:} Tears of a betrayed lover mixed with salt from seven seas. \\
\textbf{Effect:} Curse those who wronged you. The target suffers \textbf{--2 dice} to maritime/weather rolls for one session. At sea, they must roll Spirit+Resolve (DV 4) each day or suffer Harm 1 (Weather). Create a 6-segment \textit{Mother's Ire} timer. \\
\textbf{Push It:} The curse spreads to the target's allies or family; mark \textbf{2 SB (Diamonds)}. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Session.

\subsubsection*{Cantors \& Cults of Raéyn}

\textbf{The Cantor's Song.} A Cantor of Raéyn sings in \textit{shanties of the turning tide}---melodies that rise and fall with the rhythm of waves, harmonies that mimic the creak of rigging and the cry of gulls. Their voice is used to calm a restless sea, to call a favorable wind, or to sing a drowned sailor's soul back to shore. Raéyn's Cantors are often found aboard ships, their songs becoming part of the vessel's luck.

\textbf{The Cult of the Salt-Washed Circle.} Raéyn's cults gather on beaches at low tide, in sea-caves, and aboard ships during storms. The Salt-Washed Circle is a fellowship of sailors, pilots, fisherfolk, and coastal dwellers who honor the sea's gifts and respect its rages. A ritual of the Circle involves each member pouring a libation of fresh water into the sea (to honor Raéyn's tears for Khemesh) while speaking a wish for safe passage or a warning of danger ahead. Their creed is carved into driftwood and left on the tide: \textit{``The sea gives, the sea takes. We who love her learn to swim.''} Outsiders are welcome to join a libation, but they must bring salt from their home shore.

\subsubsection*{Witchcraft of Raéyn (Lay / Maritime)}

A witch who walks with Raéyn is a \textit{net-mender} or \textit{tide-seer}---one who crafts with the materials of the tide: a hull carved from wood seasoned by sea-salt, a sail woven with a prayer for fair wind, or a net tied with knots that remember every fish they have caught. In Aeler, a True-Mason might carve a ``tide stone'' at a harbor entrance, but a witch of Raéyn would instead tie a salt-twine knot around a piling, so that the current remembers to favor outgoing vessels.

\textbf{The Witch's Tool: The Wave-Seer's Shell.} A large conch shell, its inner spiral polished to a mirror sheen, rimmed with brass. Once per session, the witch may hold the shell to their ear and listen to the ``tide's memory.'' For the next scene, they gain \textbf{+2 dice} to navigation or weather-reading rolls, and may ask the GM one yes/no question about hidden currents, reefs, or approaching storms (the shell whispers the answer in a voice like waves). The shell can also be used to ``mark'' a threshold: placing it on a ship's prow or a harbor gate grants +1 die to all rolls made to defend that vessel or harbor until the next dawn.

\textbf{A Hedge Gift: Salt-Kissed Rope (Basic, 4 XP).} Once per scene, the witch may tie a simple knot in a rope, cord, or strand of hair while speaking a short prayer to Raéyn. The knot becomes a \textit{tide-knot}---when the witch unties it (as a free action), the nearest body of water (a puddle, a river, the sea) briefly rises or falls by a few inches, as if a small tide has turned. This can unstick a boat, wet a fire, or startle a guard---but never causes direct harm.

\subsubsection*{Corruption of Raéyn}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Sailor's Instinct:} +1 die to Navigation and maritime survival rolls. & \textbf{Sea-Sick on Land:} --1 die to rolls requiring prolonged time away from water or open spaces. \\
2 & \textbf{Tide's Favor:} Once/scene, treat a failed maritime or weather-related roll as a success, but mark 1 SB (Clubs). & \textbf{Weather Sensitive:} Must make a Resolve test (DV 3) to resist acting on every weather change or maritime opportunity. \\
3 & \textbf{Storm's Child:} +2 dice to rolls involving weather manipulation or sea-based combat. & \textbf{Tempestuous Nature:} Suffer 1 Fatigue when denied access to open water or weather changes for extended periods. \\
4 & \textbf{Mother's Protection:} Once/session, call upon Raéyn's favor to protect a vessel or coastal area from harm. & \textbf{Protective Instinct:} --1 die to rolls involving personal safety when others are endangered nearby. \\
5 & \textbf{Wave-Rider:} Once/session, navigate any water-based obstacle or hazard with perfect ease. & \textbf{Restless Spirit:} --1 die to rolls requiring patience or stillness; the call of the sea is constant. \\
6+ & \textbf{Storm Queen:} Once/session, become the living embodiment of the sea's power. For one scene, control weather and tides within Near range. & \textbf{Mother's Burden:} Mark +2 Obligation when using this power; the weight of protecting all who sail becomes overwhelming, causing Harm 1 (Stress) until the storm passes. \\
  &  & \textbf{Khemesh's Attention:} The Kraken's gaze turns your way; the GM may introduce oceanic complications (rogue waves, crushing pressure) as narrative fallout. \\
\end{tblr}

\subsection*{Playstyle Notes}
Raéyn is the patron of sailors, navigators, storm-callers, and anyone who lives by the sea's rhythm. Her followers gain power over tides, currents, and weather, but pay with a restless spirit and a fierce protectiveness that can turn to wrath. The rivalry with her son Khemesh is ever-present---Raéyn's gifts bring change and opportunity, but the deep always waits. Ideal for characters who love the sea, embrace change, and understand that the tide that lifts you can also drag you down.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Tides and Change:} power over currents, weather, and the rhythm of the sea
  \item \textbf{Maritime Syncretism:} a goddess of creole Kahfagia, blending mother, warrior, and storm
  \item \textbf{Storm and Calm:} the duality of protection and wrath
  \item \textbf{Motherhood and Loss:} fierce love for sailors, grief for Khemesh
  \item \textbf{Khemesh's Shadow:} the Kraken is always beneath the waves, watching
\end{itemize}

% --- Monastic Tradition: Storm Pilot (Raéyn) ---
\begin{tcolorbox}[colback=black!5,colframe=blue!50!black,title=\textbf{Storm Pilot (Raéyn)},breakable]
  \emph{The sea does not hate you. It tests you. Pass, and the waves remember your name.}

  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Wits 3+, Spirit 2+, Rite of the Wave-Tide or Cantor of Raéyn.

  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} The sea's debts are tides, not chains. Monks of the Storm Pilot tradition gain \textbf{+1 die} to resist any magical debt or geas that would bind them to a fixed port or restrict their freedom to sail. If they succeed, the contract's ink runs like water and the obligation washes away.

  \vspace{2mm}
  \noindent\textbf{Techniques (purchased as Talents):}

  \begin{tblr}{
    colspec = {X[0.7,l] X[1.8,l] X[1.1,l] X[1.4,l]},
    rowhead = 1,
    row{odd}={bg=gray!10},
    row{1}={bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Tide-Reader} (Basic, 4 XP) & Once per voyage, you may reroll a failed navigation check at sea. & 1 Boon & Wits 3 \\
  \textbf{Storm Glass} (Basic, 5 XP) & You can predict the weather for the next hour with perfect accuracy (no roll). & Passive & Spirit 2, Lore 1 \\
  \textbf{Sea Legs} (Basic, 4 XP) & You ignore all penalties from rough seas or slippery decks. & Passive & Body 2 \\
  \textbf{Call the Breeze} (Advanced, 8 XP) & Once per scene, increase your vessel's speed by one band for one exchange. & 1 Fatigue & Tier II \\
  \textbf{Whirlpool Step} (Advanced, 8 XP) & When on a ship or dock, you may move as if on solid ground even during a storm or while the vessel is heeling. & Passive & Wits 3, Athletics 2 \\
  \textbf{Kraken's Warning} (Advanced, 9 XP) & Once per session, when a sea creature (or Khemesh's influence) would ambush you, you sense it one round in advance. Your party cannot be surprised. & Free & Tier III, Spirit 3 \\
  \end{tblr}

  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Eye of the Storm.} Once per session, you may calm a supernatural storm or redirect a natural one for one scene. All weather-based penalties are negated for your vessel and allies within Near. After the scene, you gain a permanent \textbf{Storm-Touched Eye} (one iris glimmers with distant lightning, and the sea will occasionally call you to calm another tempest). You may ignore one Obligation to Raéyn, but the next natural storm you encounter will find you regardless of shelter.
\end{tcolorbox}