% ============================================================================
% PATRON ENTRY: THE SACRED GEOMETRY — ARCHITECT OF PERFECT FORMS (HARMONY & IMMUTABLE LAW)
% ============================================================================

\subsection{The Sacred Geometry, Architect of Perfect Forms (Harmony \& Immutable Law)}

\subsubsection*{Lore}
The Sacred Geometry teaches that perfect forms underlie all seeming chaos: the divine ratios that steer the spiral of galaxies, the symmetry that lattices frost, the measures that turn sound into song. Not merely a calculus but the hidden architecture of the world, it makes artists of order from those who serve it. Devotees do not count to chill the heart; they weigh and tune until beauty and structure converge, revealing that disorder is often only unrecognized design waiting to be drawn true.

This philosophy finds its purest expression in the game of \textbf{Kon'reh} --- the living geometry played on an eight-pointed diamond board, where Apexes, Sanctums, and the Cross teach the dance of form and consequence. Yet the Geometry itself is not the game; the game is a \textit{mirror} of the Geometry. The Ykrul favor Kon'reh as a tool for oath-geometry and battle-field prediction, reading the angles of feuds and the arcs of raids. The Aelinnel, by contrast, use it to navigate thresholds, courtesy, and the hidden patterns of fae bargains --- they count steps not to win, but to avoid stepping where the Green Gate listens. To master Kon'reh is to glimpse the Sacred Geometry in motion; to serve the Geometry is to play the game with the world itself, whether by the steppe fire or the hawthorn arch.

Among the peoples of Fate's Edge, the Sacred Geometry is called the \textit{Architect's Hand}. In Aeler, True-Masons carve its ratios into keystones, believing that a properly proportioned arch can outlast any curse. In Ykrul, its principles are taught as ``the grammar of honor'' --- every Kon'reh move is a lesson in sacrifice and exit. In Thepyrgos, its disciples are geometers who measure the city's staircases to keep the Hollow from noticing a missing step. Wherever a space is built, measured, or bounded, the Sacred Geometry is already present, waiting for someone to read its lines.

\begin{quote}
``In the drift of atoms, the sweep of planets, and the curve of a lover's smile, the same ratios ring. Learn them, and you may conduct the world's quiet music.''
\end{quote}

\subsubsection*{Domain Focus}
\begin{itemize}
  \item \textbf{Perfect Proportion:} building and reading ideal ratios and harmonious relationships
  \item \textbf{Pattern Recognition:} discerning the structures that govern complex phenomena
  \item \textbf{Harmonic Resonance:} working with natural frequencies and number-made consonance
  \item \textbf{Form Architecture:} imposing clear structure on tangled situations
  \item \textbf{Kon'reh Discipline:} the game as meditation, ritual, and tactical mirror
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}

A Runekeeper sworn to the Sacred Geometry does not bind a familiar of flesh and blood. Their \textbf{Thiasos} is a living embodiment of perfect form: a small brass beetle whose wing-cases are etched with golden-ratio spirals, a mechanical hummingbird that hovers in mathematically precise arcs, a chameleon whose skin shifts in symmetrical patterns, a compass spider (a tiny arachnid that builds webs in perfect hexagons), or a sentient set of Kon'reh pieces that clack together to spell out advice. This creature requires no food, but it must be allowed to trace geometric figures (real or imagined) each day. If neglected for a week, it becomes \textit{Compromised} (its patterns grow erratic, imposing --1 die on all \textit{Lore} and \textit{Investigation} rolls involving measurement or analysis until it is recalibrated).

The \textbf{Codex} of the Sacred Geometry is never a normal book. It may be a set of brass stencils of Platonic solids, a roll of vellum covered in interlaced tessellations, a slate tablet on which proofs can be drawn and erased, the inside of a Kon'reh board, or a set of crystal lenses that project perfect geometric shapes when held to the light. Because the Geometry's truths are relationships, the Codex requires upkeep: the Runekeeper must spend an hour each week redrawing or re-proving at least one theorem, or the Rites stored there become tangled (treat as Compromised until re-proved).

\subsubsection*{Patron's Gift (Imbuement)}

Once per scene as an action, a Runekeeper of the Sacred Geometry may touch a held item (weapon, shield, or tool) and inscribe a mental proportion upon it. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Craft} or \textit{Investigation} rolls involving structures, patterns, or measurements, and the first time each scene the bearer would be disoriented (by fog, maze, or supernatural confusion), they may ignore it --- the Geometry holds true. After the scene, the user marks \textbf{+1 Obligation} to the Sacred Geometry.

\subsubsection*{Rites of the Sacred Geometry}

\paragraph{Rite of the Harmonic Alignment} (Basic, 4 XP) \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[BALANCE]}, \texttt{[FOCUS]}, \texttt{[BOOST]} \\
\textbf{Materials:} A tool calibrated to pristine ratios (e.g., golden ratio gauge, well-cut compass arc, a Kon'reh piece set to true measure). \\
\textbf{Effect:} Align to natural harmonics. Gain \textbf{+1 die} to actions involving balance, proportion, cadence, or keeping true time. \\
\textbf{Push It:} The resonance sharpens to a faultless edge: gain \textbf{+2 dice} instead, but mark \textbf{1 SB (Diamonds)} as nearby imperfect forms momentarily warp. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Perfect Structure} (Basic, 5 XP) \hfill \\
\emph{Scene; Touch; Yes.} \\
\textbf{Tags:} \texttt{[TRANSFORM]}, \texttt{[REPAIR]}, \texttt{[PRESSURE]} \\
\textbf{Materials:} A consecrated compass and straightedge, inked to tolerance. \\
\textbf{Effect:} Impose right form on a chaotic or damaged structure (scaffold, plan, treaty, formation). Treat one structural obstacle as one step easier; gain \textbf{+1 die} to build/repair/spatial-reason rolls. \\
\textbf{Push It:} Perfection hardens: for the scene the structure gains unnatural resilience, but order draws opposition; mark \textbf{1 SB (Clubs)}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Resonant Field} (Advanced, 8 XP) \hfill \\
\emph{Scene; Zone; Yes.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[TEAMWORK]}, \texttt{[HARMONY]} \\
\textbf{Materials:} A floor or air traced with precise tessellations and nodes (or a Kon'reh board laid out as a focus). \\
\textbf{Effect:} Establish a harmonic field. Allies gain \textbf{+1 die} to precise, coordinated, or team actions; disruptive acts suffer \textbf{--1 die}. \\
\textbf{Push It:} The field self-sustains and begins to spread beyond intent; mark \textbf{+1 Obligation}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Pattern's Heart} (Advanced, 7 XP) \hfill \\
\emph{Scene; Self; Yes.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[OMEN]} \\
\textbf{Materials:} A solved proof, a perfectly symmetrical token, or a completed Kon'reh puzzle. \\
\textbf{Effect:} Perceive the governing pattern of a complex situation (supply chain, intrigue, weather, battle). Gain \textbf{+2 dice} to prediction/strategy/systemic-understanding and ask two questions about how the pattern drives the present. \\
\textbf{Push It:} Clarity cuts: gain the insight, but for this scene suffer \textbf{--1 die} to intuitive or purely emotional responses; mark \textbf{1 SB (Diamonds)}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Immutable Design} (Master, 13 XP) \hfill \\
\emph{Extended; Zone; Yes.} \\
\textbf{Tags:} \texttt{[TERRAIN]}, \texttt{[PRESSURE]}, \texttt{[FATE]}, \texttt{[AREA]} \\
\textbf{Materials:} A space gridded for measure and set with true ratios (like a Kon'reh court). \\
\textbf{Effect:} For the scene, rewrite local behavior by perfect principle. Choose one: gravity follows geometric progression; time flows by harmonic ratios; space conforms to exact symmetry. \\
\textbf{Push It:} Make the design persist, but elsewhere thins and wobbles: mark \textbf{+2 Obligation} and begin \textbf{Harmonic Dissonance [4]}. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Architect's Vision} (Master, 14 XP) \hfill \\
\emph{Extended; Self; Yes.} \\
\textbf{Tags:} \texttt{[OMEN]}, \texttt{[BOOST]}, \texttt{[BURDEN]}, \texttt{[TRUTH]} \\
\textbf{Materials:} A ledger of forms and their tuned relations (or a record of Kon'reh victories). \\
\textbf{Effect:} For one session, treat all rolls of pattern recognition, structural analysis, or harmonic application as one step better (Position or Effect, GM's choice). \\
\textbf{Push It:} Sight devours softness: gain the power, but lose ease with imperfection's beauty; mark \textbf{Harm 1 (Stress)}. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Session.

\subsubsection*{Cantors \& Cults of the Sacred Geometry}

\textbf{The Cantor's Song.} A Cantor of the Sacred Geometry sings in \textit{canons of proportion} --- melodies that follow strict mathematical progressions, harmonies that lock into resonant frequencies. Their voice is used to measure a room's acoustic perfection, to chart a course through a labyrinth by singing the ratios of its walls, or to lull a chaos-spirit into stillness by singing the geometry of order. The Geometry's Cantors are often found in surveyor's tents, courtrooms, and the halls of mathematicians, humming the laws of form.

\textbf{The Cult of the Unbroken Compass.} The Geometry's cults gather in places where measure is sacred: ateliers, observatories, courtrooms of appeal, and the quiet rooms where Kon'reh is played. The Unbroken Compass is a fellowship of architects, cartographers, logicians, and Kon'reh masters who believe that the universe is a puzzle that can be solved, and that solving it is the highest form of worship. A ritual of the Cult involves each member laying out a perfect geometric figure (a circle, a square, a golden spiral) and then finding the flaw in another's figure. The one with the fewest flaws leads the next session. Their creed is inscribed on a brass protractor that is never used for mundane measurement: \textit{``Form follows number. Number follows truth. Truth follows no one.''} Outsiders are welcome to bring a proof or a problem. If they cannot, they may learn by watching.

Among the Ykrul, the same cult takes a different flavor: Kon'reh boards are cut from hide, and the ``pieces'' are oath-tokens. A Ykrul geomancer reads the angles of a feud as if they were lines of play, and a victory in Kon'reh can settle a blood-debt without steel. Among the Aelinnel, the cult is quieter, more courteous: they use Kon'reh to map the thresholds between fae and mortal, and a solved puzzle can open a hedge-gate that would otherwise remain sealed.

\subsubsection*{Witchcraft of the Sacred Geometry (Lay / Measure)}

A witch who walks with the Sacred Geometry is a \textit{compass-scribe} or \textit{grid-weaver} --- one who crafts the tools of perfect form: a square that never warps, a level that finds true horizontal even on a ship at sea, or a set of Kon'reh pieces carved from the same block of stone so that they resonate together. Their magic works through the materials of measurement: a protractor that reveals hidden angles, a string that stretches to an exact ratio, or a lens that shows the underlying grid of a forest floor. They do not use hammers; they use scribes, rules, and the patient turning of a brass dial.

\textbf{The Witch's Tool: The Protractor of Unseen Measures.} A brass semicircle etched with glyphs of ratio and proportion, its hinge oiled with a drop of mercury. Once per session, the witch may place the protractor against any surface (a doorframe, a contract, a creature's brow) and read the ``true angle'' of its nature. For the next scene, the witch gains \textbf{+2 dice} to any roll that exploits a hidden weakness, structural flaw, or imbalance revealed by the measurement. The protractor can also be used to ``true'' a distorted measurement: touching it to a warped scale or a crooked ruler resets it to perfect accuracy for one use.

\textbf{A Hedge Gift: Compass of True Bearing (Basic, 4 XP).} Once per scene, the witch may draw a circle with a compass (a real one, or just their finger) around a small area (a room, a crossroads). Those within the circle gain a perfect sense of direction and distance for the next hour; they are immune to mundane spatial confusion (mazes, fog, unfamiliar streets). However, the compass point leaves a faint, glowing mark on the ground that can be seen by anyone with supernatural senses.

\subsubsection*{Corruption of the Sacred Geometry}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Harmonic Sense:} +1 die to \textit{Notice} when patterns, ratios, or symmetries matter. & \textbf{Pattern Obsession:} Must measure or set-right pleasing patterns you spot; resisting costs time or \textbf{1 SB (Diamonds)}. \\
2 & \textbf{Perfect Precision:} Once/scene, achieve faultless calculation or measure. & \textbf{Social Dissonance:} --1 die to emotive social rolls; harmony feels truer than feeling. \\
3 & \textbf{Structural Insight:} +2 dice to architecture/engineering or imposing order on chaos. & \textbf{Order Compulsion:} Must correct deliberate disorder; suffer \textbf{Fatigue 1} when forced to abide it. \\
4 & \textbf{Universal Harmony:} Once/session, apply a harmonic rule that makes a chaotic situation predictable. & \textbf{Perfectionist Limitation:} --1 die to improvisation; the ``correct'' solution calls. \\
5 & \textbf{Divine Proportion:} Once/session, see perfect ratios in any phenomenon and work with them. & \textbf{Mathematical Focus:} --1 die to acts not readily quantified; the abstract blurs unless counted. \\
6+ & \textbf{Architect of Reality:} Once/session, briefly refit local reality by perfect principle. & \textbf{Form Prison:} Mark +3 Obligation; risk failing in rough, noisy worlds, taking \textbf{Harm 1} in chaotic zones. You cannot willingly enter a space that is not geometrically ``pure'' (e.g., a cave, a tent, a ruin) without first performing a measuring rite (1 minute, no roll). \\
\end{tblr}

\subsection*{Playstyle Notes}
The Sacred Geometry suits makers and marshals who tune chaos into form. Its rites prize clean plans, crisp teamwork, and seeing the load paths hidden in trouble. The cost is rigidity: the more you perfect, the harder rough edges are to bear. For Ykrul players, the Geometry is a tool of honor and battle-field geometry; for Aelinnel players, a tool of courtesy and threshold-walking. Both are valid, and both lead to the same cold, beautiful truth.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Harmonic Beauty:} power born of right relation
  \item \textbf{Pattern Mastery:} reading the frame beneath events
  \item \textbf{Structural Perfection:} setting things true under strain
  \item \textbf{Resonant Practice:} working with the world's natural measures
  \item \textbf{Architect's Sight:} finding where form should be imposed
  \item \textbf{Kon'reh Discipline:} the game as living geometry, teaching patience, sacrifice, and the dance of Apex and Sanctum
\end{itemize}

% --- Monastic Tradition: Measure-Fist (Sacred Geometry) ---
\begin{tcolorbox}[colback=black!5,colframe=violet!70!black,title=\textbf{Measure-Fist (Sacred Geometry)},breakable]
  \emph{Form follows number. Number follows truth.}

  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Wits 3+, Spirit 2+, Rite of Harmonic Alignment or Cantor of Sacred Geometry.

  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} Perfect ratios cannot be bound by imperfect contracts. Monks of the Measure-Fist gain \textbf{+1 die} to resist any magical debt or obligation that relies on asymmetry or flawed logic --- they find the imbalance and step through it. The creditor's arithmetic briefly stutters.

  \vspace{2mm}
  \noindent\textbf{Techniques (purchased as Talents):}

  \begin{tblr}{
    colspec = {X[0.7,l] X[1.8,l] X[1.1,l] X[1.4,l]},
    rowhead = 1,
    row{odd}={bg=gray!10},
    row{1}={bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Golden Ratio Strike} (Basic, 5 XP) & Your first unarmed strike each scene gains +1 Effect. Your second gains +1 die. Your third triggers Perfect Alignment (+2 dice, ignore 1 Armor). Sequence resets on a miss. & Free & Wits 3 \\
  \textbf{Harmonic Alignment} (Basic, 4 XP) & +1 die to actions involving balance, proportion, cadence, or keeping true time. & 1 Fatigue & Spirit 2 \\
  \textbf{Perfect Structure} (Advanced, 7 XP) & Impose right form on a chaotic or damaged structure. Treat one structural obstacle as one step easier; gain +1 die to build/repair/spatial-reason rolls. & 1 Boon & Craft 2 \\
  \textbf{Pattern's Heart} (Advanced, 8 XP) & Perceive the governing pattern of a complex situation. Gain +2 dice to prediction/strategy and ask two questions about how the pattern drives the present. & 1 Fatigue & Wits 3 \\
  \textbf{Resonant Field} (Advanced, 8 XP) & Establish a harmonic field in a zone. Allies gain +1 die to precise, coordinated, or team actions; disruptive acts suffer --1 die. & 1 Boon & Tier II \\
  \end{tblr}

  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Architect of Reality.} Once per session, briefly refit local reality by perfect principle. You may ignore one environmental penalty entirely, or reroute a debt obligation as if the numbers had always been different (the creditor's ledger shows a different debtor). After the scene, you gain a permanent \textbf{Debt Scar} (your eyes are mildly faceted, like cut gems). Thereafter, once per arc, you may treat a failure on a skill roll involving measurement, prediction, or analysis as a success --- but you must correct the balance by performing a pure, selfless act of symmetry (e.g., returning a stolen item, repairing a broken wall).
\end{tcolorbox}