% --- Patron: The Savage Heart (Red Maw) ---
\subsection{The Savage Heart (Red Maw)}

\textit{Lore.} The Savage Heart is the truth behind the green veil: nature not as garden, but as gullet. It is the claw that tears, the fang that crushes, the hunger that stalks without malice and kills without mercy. Its followers are the druids you do not meet by firelight — those who have learned that the forest feeds on its own, that the strong devour the weak, and that the only law is the one written in blood on snow.

Across the known world, the Savage Heart wears the face of the apex predator. In the Amaranthine core, it is the \textit{Great Bear}, patient and unstoppable, a wall of fur and fury. Among the Ykrul and Kuvani, it is the \textit{Wolf of the Steppe}, a pack hunter with tireless stamina. In Oshiira, it is the \textit{Golden Lion}, whose roar shatters the hunt's silence. In Sekogo, it is the \textit{Leopard of the Tall Grass}, a shadow that strikes once and is gone. In Dhahara, it is the \textit{Tiger of the Mangroves}, whose amber gaze promises a death that runs faster than you can flee. Wherever prey fears the dark, the Savage Heart is already there, waiting.

\begin{quote}
``The deer does not weep for the wolf. The sapling does not beg the axe. In the red maw, there is no cruelty — only truth. The truth that all flesh feeds the earth, and the earth feeds the strong.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Apex Predation:} hunting, stalking, dominating the food chain
  \item \textbf{Red Tooth:} killing as nature's sacrament, strength proven through violence
  \item \textbf{Scavenger's Wisdom:} using death for growth, nothing wasted
  \item \textbf{Territory and Pack:} dominance over a domain, leadership of the hunt
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to the Savage Heart does not bind a familiar of hearth and home. Their \textbf{Thiasos} is a small predator or hunting companion: a wolf pup with eyes the color of amber, a lynx kitten whose fur prickles at danger, a hawk with one white feather that never misses its mark, a viper that coils around the wrist and tastes the air for fear, or a shadow-fox that follows silently and leaves no tracks. This creature requires no food in the usual sense, but it must hunt and kill its own prey each week (a rat, a rabbit, or larger, depending on the keeper's means). If neglected, it becomes \textit{Compromised} (it grows listless, and its keeper loses the benefit of \textit{Stalker's Instinct} until the creature is fed with a fresh kill).

The \textbf{Codex} of the Savage Heart is never a book of paper and ink. It may be a belt of predator teeth, each tooth carved with a hunting verse; a bundle of cured hides painted with the kill-records of the keeper's lineage; a set of bone tally sticks, notched for every successful hunt; or the skull of a great beast, its jaw etched with runes that can only be read by the light of a fresh kill. Because the hunt resists being fixed in writing, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week oiling the leather, sharpening the teeth, or adding a new notch for a recent kill; otherwise the Rites stored there fade (treat as Compromised until the keeper makes an offering of fresh blood).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of the Savage Heart may touch a held item (weapon, tool, or piece of armor) and whisper a hunter's promise. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Athletics} or \textit{Stealth} rolls when in natural terrain, and the first time each scene the bearer would be knocked prone, they may ignore it (the hunter's stance is sure). After the scene, the user marks \textbf{+1 Obligation} as the scent of the hunt clings to them.

\subsubsection*{Rites of the Savage Heart}

\paragraph{Rite of the Predator's Gaze (Low, 4 XP)} \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[MARK]}, \texttt{[FEAR]}, \texttt{[TRACK]} \\
\textbf{Materials:} A tooth from a predator that killed its own prey. \\
\textbf{Effect:} Mark a single target as your prey. Gain \textbf{+1 die} on all rolls to track, pursue, or intimidate that target. The effect lasts until the target escapes or is killed. \\
\textbf{Push It:} The prey feels your gaze; they suffer --1 die to resist social pressure from you. Mark \textbf{1 SB (Spades)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Blood Scent (Low, 5 XP)} \hfill \\
\emph{Scene; Self; Yes.} \\
\textbf{Tags:} \texttt{[SENSE]}, \texttt{[FEAR]}, \texttt{[TRACK]} \\
\textbf{Materials:} A fresh kill's blood smeared on your face. \\
\textbf{Effect:} You scent the air like a beast. Gain \textbf{+2 dice} to \textit{Investigation} or \textit{Survival} when tracking by smell. You may also detect fear — you sense if a creature within Near is Frightened. \\
\textbf{Push It:} Your sense expands to Far, but you are overwhelmed by the smell of blood; mark \textbf{1 SB (Hearts)} and suffer --1 die to social rolls until the end of the scene. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Shifting Shape (Standard, 8 XP)} \hfill \\
\emph{Scene; Self; Yes.} \\
\textbf{Tags:} \texttt{[TRANSFORM]}, \texttt{[BOOST]} \\
\textbf{Materials:} A pelt or mask from the beast whose form you take. \\
\textbf{Effect:} Assume the form of a predator for the scene. Choose one:

\medskip
\noindent\textbf{Bear (Tank Form)}
\begin{itemize}
  \item \textbf{Passive:} +1 Armor, +2 dice to Body rolls.
  \item \textbf{Guardian's Pounce:} Once per scene, when you protect an ally, you may also knock the attacker prone (Body+Resolve DV 3 or fall).
  \item \textbf{Limitation:} Movement reduced by one band (cannot Dash).
\end{itemize}

\noindent\textbf{Feline or Canine (Striker Form)}
\begin{itemize}
  \item \textbf{Passive:} +2 dice to Agility and Stealth, +1 band to movement (can Dash as a free action once per scene).
  \item \textbf{Ambush Predator:} When attacking from stealth, gain +1 Effect and ignore 1 point of the target's Armor.
  \item \textbf{Limitation:} --1 Armor (or +1 Harm taken).
\end{itemize}

Cultural variations: Bear/Wolf (north), Lion/Leopard (Oshiira), Tiger/Panther (Dhahara). You may speak but cannot use tools or cast other Rites while transformed. \\
\textbf{Push It:} Shift as a free action once this scene, but mark \textbf{1 SB (Clubs)} as the beast presses against your mind. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Pack Hunt (Standard, 7 XP)} \hfill \\
\emph{Scene; Near; Yes.} \\
\textbf{Tags:} \texttt{[TEAMWORK]}, \texttt{[TRANSFORM]}, \texttt{[BOOST]} \\
\textbf{Materials:} A token from each member of your pack (a claw, a feather, a scrap of hide). \\
\textbf{Effect:} You and up to three allies form a hunting pack. While within Near of each other, all gain \textbf{+1 die} to attacks against a single designated prey, and the first ally to attack each round gains \textbf{+1 Effect}. \\
\textbf{Push It:} Coordination extends to Far, but hunters become territorial; mark \textbf{1 SB (Spades)} as rivalries flare. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Carrion Feast (Standard, 9 XP)} \hfill \\
\emph{Scene; Touch; Yes.} \\
\textbf{Tags:} \texttt{[CONSUME]}, \texttt{[HEAL]}, \texttt{[BOOST]} \\
\textbf{Materials:} The flesh of a fallen enemy, consumed raw. \\
\textbf{Effect:} Devour the strength of the fallen. Heal \textbf{1 Fatigue} immediately, and gain \textbf{+1 die} to your next physical action. If the enemy was supernatural (spirit, demon, fey), you also gain a minor property of theirs for one scene. \\
\textbf{Push It:} Heal 1 Harm instead of Fatigue, but mark \textbf{1 SB (Hearts)} as the feast attracts scavengers. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Final Hunt (High, 14 XP)} \hfill \\
\emph{Extended; Near; No.} \\
\textbf{Tags:} \texttt{[TRANSFORM]}, \texttt{[FEAR]}, \texttt{[SPEED]}, \texttt{[DEATH]} \\
\textbf{Materials:} The heart of a predator you have killed in single combat, still warm. \\
\textbf{Effect:} You become the ultimate hunter for one scene:
\begin{itemize}
  \item Ignore all difficult terrain.
  \item Gain \textbf{+3 dice} to tracking and combat rolls against your designated prey.
  \item If you reduce a prey to 0 Harm, make a free attack against another target within Near.
\end{itemize}
\textbf{Push It:} The hunt consumes you — you cannot end the rite until all designated prey are dead or fled, and you suffer Harm 1 (Weariness) when the scene ends. Mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 8 segments.

\subsubsection*{Cantors \& Cults of the Savage Heart}

\textbf{The Cantor's Song.} A Cantor of the Savage Heart sings in \textit{hunting calls} — short, sharp cries that mimic the bark of a bear or the cough of a leopard. Their voice is used to coordinate a pack, to scare prey from hiding, or to summon a predator from the deep woods. They are rarely heard near villages; their songs echo in the places between game trails.

\textbf{The Cult of the Painted Maw.} The Savage Heart's cults do not meet in groves — they gather at kills, wearing masks of bone and hide that show the animal they have sworn to become. The most ancient sect, the \textit{Painted Maw}, ritually scarifies each member's chest with the claw marks of their totem beast; the scars are filled with ash and charcoal, and the wounds are licked clean by the pack's eldest hunter. A ritual involves each member speaking the name of a kill they made alone, then adding a tooth from that kill to a communal jawbone that hangs from a tree. Their creed is spoken with full mouths: \textit{``The forest is a stomach. We are its teeth.''} Outsiders are not welcome; they are either prey or future members. Those who join must prove themselves by returning with the heart of a predator they have killed with no weapon but their own hands.

\subsubsection*{Witchcraft of the Savage Heart (Lay / Hunter)}

A witch who walks with the Savage Heart is a \textit{skin-stitcher} or \textit{tooth-carver} — but their tools are never forge-hammers. They work with bone, claw, fur, and sinew, crafting masks that hold the spirit of a hunted beast, or drums of leopard skin that call the hunt. Their magic is the patience of the stalker and the savagery of the kill. In Dhahara, a tiger-witch might sew a cloak from the hide of a man-eater, wearing it to walk unseen through the mangroves. In the Ykrul steppe, a wolf-witch might carve a whistle from a wolf's femur that, when blown, answers only with the howl of the pack.

\textbf{The Witch's Tool: The Hunter's Mask.} A mask carved from the skull of a predator, lined with fur and painted with the blood of the wearer's first kill. Once per session, when you don the mask, you gain the senses of that predator for the scene: \textbf{+2 dice} to tracking by scent or sound, and you may ask the GM one yes/no question about a creature's fear or weakness. The mask also allows you to speak with animals of that species (they will answer simple questions, but they are never safe to trust).

\textbf{A Hedge Gift: Hunter's Snuff (Basic, 4 XP).} Once per scene, the witch may blow a pinch of powdered bone into the air. For the next minute, the witch can smell the fear of any creature within Near range — they know if a target is Frightened, and gain \textbf{+1 die} to track it.

\subsubsection*{The Savage Heart's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1.5cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Stalker's Instinct:} +1 die to tracking and stealth in wilderness. & \textbf{Blood Hunger:} Must kill or consume fresh meat each day or suffer 1 Fatigue. \\
2 & \textbf{Apex Demeanor:} +1 die to Intimidation against those weaker than you. & \textbf{Pack Guilt:} --1 die to social rolls in civilized company; you smell faintly of carrion. \\
3 & \textbf{Scavenger's Resilience:} +2 dice to resist disease and poison from raw meat. & \textbf{Compulsive Dominance:} Must assert physical dominance when challenged; --1 die to de-escalation. \\
4 & \textbf{Wound Licker:} Once/session, heal 1 Fatigue by licking a fresh wound (your own or ally's). & \textbf{Beastly Tells:} Your teeth sharpen, your eyes reflect light; --1 die to disguise. \\
5 & \textbf{Alpha's Roar:} Once/session, make a terrifying sound that forces all nearby prey to test Resolve (DV 4) or flee. & \textbf{Lone Wolf:} --2 dice to cooperate with those who have not proven themselves in battle. \\
6+ & \textbf{Maw Incarnate:} Once/session, become a true predator. For one scene, all attacks ignore 1 Armor and you heal 1 Fatigue on each kill. & \textbf{The Red Eats All:} Mark +3 Obligation; risk becoming feral, unable to speak or reason until you have fed again. \\
\end{tblr}

\subsection*{Playstyle Notes}
The Savage Heart rewards those who hunt, kill, and dominate. Its followers are not gentle guardians of nature — they are its teeth and claws. Power comes from taking, from proving strength, from the raw act of predation. The cost is the slow loss of the civilized self, the creeping realization that you are becoming the thing you stalk. The Bear form excels at absorbing damage and protecting allies; the Feline or Canine form excels at striking from stealth and mobility. Neither is strictly better than a dedicated specialist — a true tank has more armor, a true striker deals more burst damage — but the Savage Heart's gift is \textit{versatility within the hunt}.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Predation as Sacrament:} killing is holy when done with purpose
  \item \textbf{Territory and Dominance:} claim, defend, expand
  \item \textbf{Pack Tactics:} hunting is a team sport
  \item \textbf{Red Tooth Wisdom:} death feeds life, cruelty is not malice but nature
  \item \textbf{Transformation:} becoming the beast to master the beast
\end{itemize}

\subsection*{Rivalries}
\begin{itemize}
  \item \textbf{Grimmir (Seasonal Wisdom):} Grimmir preserves the balance; the Savage Heart devours it. They are two faces of the same coin — but the coin is always flipped.
  \item \textbf{The Sealed Gate (Containment):} The Gate wards things out; the Savage Heart tears down walls.
  \item \textbf{The Sacred Geometry (Harmony):} Order and perfect form are prey to the red maw.
\end{itemize}

\subsection*{Obligation Triggers}
\begin{itemize}
  \item \textbf{8+ Segments:} The Savage Heart demands you take a worthy trophy — hunt a creature of equal or greater strength than yourself, alone.
  \item \textbf{10+ Segments:} You hear the call of the hunt always. Refusing to answer when prey is weak will cost you the use of your Rites until you kill again.
\end{itemize}

\subsection*{Signs of Favor}
Your eyes reflect light in darkness, your shadow seems to move independently, smaller animals flee your presence, and the scent of blood often follows you (real or imagined).

% --- Monastic Tradition: Beast-Form (Savage Heart) ---
\begin{tcolorbox}[colback=black!5,colframe=brown!70!black,title=\textbf{Beast-Form (Savage Heart)}]
  \emph{The forest is a stomach. We are its teeth.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Body 3+, Spirit 2+, Rite of the Predator's Gaze or Cantor of Savage Heart.
  
  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} The Maw does not lend — it devours. Monks of the Savage Heart gain \textbf{+1 die} to resist any magical debt or obligation that would compel them to act against their predatory nature. If they succeed, the creditor's paperwork briefly smells of fresh blood.
  
  \vspace{2mm}
  \noindent\textbf{Techniques (purchased as Talents):}
  
  \begin{tblr}{
    colspec = {X[0.7,l] X[1.8,l] X[1.1,l] X[1.3,l]},
    rowhead = 1, row{odd}={bg=gray!10}, row{1}={bg=gray!30}, hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Predator's Stance} (Basic, 5 XP) & Choose Bear (Tank: +1 Armor, +2 dice Body) or Feline (Striker: +2 dice Agility/Stealth, +1 band movement). Lasts one scene. & 1 Fatigue & Body 3 \\
  \textbf{Blood Scent} (Basic, 5 XP) & Gain +2 dice to track by smell. Detect fear — sense if a creature within Near is Frightened. & 1 Boon & Spirit 2 \\
  \textbf{Scavenger's Resilience} (Basic, 4 XP) & +2 dice to resist poison or disease from raw meat. & Passive & Body 3 \\
  \textbf{Pack Hunter} (Standard, 7 XP) & You and up to three allies form a hunting pack. While within Near of each other, all gain +1 die to attacks against a single designated prey, and the first ally to attack each round gains +1 Effect. & 1 Fatigue & Tier II \\
  \textbf{Carrion Feast} (Standard, 9 XP) & Consume the flesh of a fallen enemy. Heal 1 Fatigue immediately, and gain +1 die on your next physical action. If the enemy was supernatural, gain a minor property of theirs for one scene. & 1 Boon & Tier II \\
  \textbf{Alpha's Roar} (Standard, 8 XP) & Once per session, make a terrifying sound that forces all nearby prey to test Resolve (DV 4) or flee. & Free & Tier III \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Maw Incarnate.} Once per session, become a true predator for one scene: all attacks ignore 1 Armor and you heal 1 Fatigue on each kill. After the scene, you take Harm 2 (Stress) and gain a permanent \textbf{Debt Scar} — a scar in the shape of a fang mark. You may ignore one debt or obligation that would force you to act against your pack (the creditor hears only growling).
\end{tcolorbox}