% --- Patron: The Silent Choir, Keeper of Unspoken Truths (Restraint & Protective Omission) ---
\subsection{The Silent Choir, Keeper of Unspoken Truths (Restraint \& Protective Omission)}

\textit{Lore.} The Silent Choir is the gathered voice of truths wisely left unsaid, confessions granted mercy, and secrets kept to spare a life from breaking. It is not absence but presence: the practiced wisdom that some knowledge wounds more than it heals. Its keepers are wardens of confidence and shelterers of the vulnerable. They learn to carry burdens alone, to let volatile truths sleep, and to shield others from fires that revelation would feed.

Unlike the Witness, who reveals truth at any cost, or Ikasha, who hides secrets for power, the Silent Choir teaches \textit{restraint as protection}. A secret is not a weapon; it is a weight. To speak it may free the speaker but shatter the listener. The Choir’s followers judge not what is true, but what is kind. They are therapists, not inquisitors; confessors, not spies.

In the world of Fate's Edge, the Silent Choir is invoked by bell‑wardens who keep a certain pattern unsung to avoid calling the fog, archivists who seal dangerous texts behind a second lock and whisper the combination to no one, and death‑doulas who carry a dying oath to the grave. The Choir has no temples, only quiet rooms, sealed letter‑boxes, and the spaces between heartbeats where a truth could be told – but is not.

Among the peoples of Fate's Edge, the Silent Choir is known as the \textit{Grey Watchers} in Thepyrgos, where bell‑wardens maintain the silence furlongs. In Aeler, their practice is called “breath‑stillness” – a meditative art used by vent‑priors to keep dangerous pressures from being spoken aloud. In Vhasia, they are the \textit{Silent Notaries}, who witness death‑bed testaments and then bury the name of the inheritor in a lead box that is never opened. Wherever a secret must be protected, the Silent Choir is already there, listening – and choosing not to speak.

\begin{quote}
``In a world that shouts, the Choir teaches the weight of quiet. Not every wound needs salt, not every blaze needs breath, not every heart can bear the full truth.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Protective Silence:} when and how to keep counsel for others’ good
  \item \textbf{Merciful Omission:} judging truths that harm when spoken
  \item \textbf{Confidential Guardianship:} safeguarding people and perilous knowledge
  \item \textbf{Restraint Wisdom:} the power of what is not said or done
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to the Silent Choir does not bind a familiar of sound and fury. Their \textbf{Thiasos} is a small creature that embodies quiet discretion: a grey mouse that nests in letter‑boxes and never cheeps, a moth whose wings are dusted with the powder of forgetfulness, a small owl that flies without stirring air, a whippoorwill whose call is a barely audible sigh, or a tiny toad that hibernates in sealed jars and wakes only when a secret is told. This creature requires no food, but it must be present when a confidence is kept – when someone refrains from speaking a harmful truth. If neglected for a week (no witnessed restraint), it becomes \textit{Compromised} (it emits a faint, anxious buzzing; its bearer gains –1 die to all rolls involving discretion or secrecy until it is soothed with a whispered promise).

The \textbf{Codex} of the Silent Choir is never a book meant to be read aloud. It may be a locked journal whose pages are blank until breathed upon, a set of sealed wax tablets that crumble if opened by the wrong hands, a collection of memory‑knots tied in grey silk, or the inside of a bell that has been silenced with felt. Because the Choir’s truths are kept in omission, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week reviewing the secrets they have carried (without speaking them aloud) or the Rites stored there become faint (treat as Compromised until re‑sealed).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of the Silent Choir may touch a held item (weapon, shield, or tool) and whisper a word of stillness. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Stealth} or \textit{Insight} rolls involving discretion or secrecy, and the first time each scene the bearer would be forced to reveal a secret by magical compulsion, they may ignore it (the Choir’s silence protects them). After the scene, the user marks \textbf{+1 Obligation} as the weight of the unspoken presses upon them.

\subsubsection*{Rites of the Silent Choir}

\paragraph{Rite of the Gentle Seal (Low, 5 XP)} \hfill \\
\emph{Scene; Near/Touch; Yes.} \\
\textbf{Tags:} \texttt{[SILENCE]}, \texttt{[PROTECT]}, \texttt{[BOOST]}, \texttt{[DEBT]} \\
\textbf{Materials:} A white feather or cloth that has drunk whispered confidences; or a sealed envelope or locked reliquary holding a sensitive fact. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Gentle Silence:} Hush a space with respectful quiet. All present gain \textbf{+1 die} to actions of discretion, confidentiality, or de‑escalation this scene.
  \item \textbf{Shielded Truth:} Ward one piece of information from discovery. Gain \textbf{+2 dice} to resist extraction or revelation of the protected truth.
\end{itemize}
\textbf{Push It:} (Shielded Truth) Protection becomes absolute but dependent: removing the ward triggers \textbf{1 SB (Hearts)} as bottled truths surge. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Unspoken Accord (Standard, 8 XP)} \hfill \\
\emph{Scene; Near; Yes.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[BOND]}, \texttt{[SILENCE]}, \texttt{[TRUTH]} \\
\textbf{Materials:} A pact signed in fading ink or a vow witnessed in silence. \\
\textbf{Effect:} Bind an understanding without paper teeth. All parties gain \textbf{+1 die} when honouring the accord; breach draws social costs even without proof. \\
\textbf{Push It:} The accord is supernaturally enforced: a breaker suffers swift ostracism; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Burden Bearer (Standard, 7 XP)} \hfill \\
\emph{Scene; Touch; Yes.} \\
\textbf{Tags:} \texttt{[TRANSFER]}, \texttt{[HEAL]}, \texttt{[DEBT]}, \texttt{[FEAR]} \\
\textbf{Materials:} An object heavy with someone's shame or grief. \\
\textbf{Effect:} Take another's social or emotional weight for a time. The target gains relief (\textbf{+2 dice} to social rolls) while you suffer the drag (\textbf{--1 die} to social rolls) this scene. \\
\textbf{Push It:} A shard remains: gain insight, keep a sliver of pain; mark \textbf{1 SB (Diamonds)}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Sanctuary of Secrets (High, 13 XP)} \hfill \\
\emph{Extended; Zone; Yes.} \\
\textbf{Tags:} \texttt{[WARD]}, \texttt{[AREA]}, \texttt{[SILENCE]}, \texttt{[PROTECT]}, \texttt{[DEBT]} \\
\textbf{Materials:} A room seasoned by trusted confidences and sealed with hush‑signs. \\
\textbf{Effect:} Consecrate a sanctuary where secrets hold fast. Truths cannot be compelled; all within gain immunity to forced revelation and \textbf{+2 dice} to maintain discretion. \\
\textbf{Push It:} The sanctuary becomes a keep: none may leave until each offers one guarded truth; mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Choir's Wisdom (High, 14 XP)} \hfill \\
\emph{Extended; Self; Yes.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[PREDICTION]}, \texttt{[FATE]}, \texttt{[BOOST]}, \texttt{[BURDEN]} \\
\textbf{Materials:} The weight of many secrets rightly kept and debts of trust well paid. \\
\textbf{Effect:} Read the branching costs of revelation before speech. Gain \textbf{+3 dice} to judge when to speak and when to keep counsel. \\
\textbf{Push It:} Sight is too wide: see all outcomes, struggle to choose; mark \textbf{Harm 1 (Stress)}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 8 segments.

% --- Confidant Keeper (Oath-Keeper) tcolorbox ---
\begin{tcolorbox}[colback=black!5,colframe=blue!70!black,title=\textbf{Confidant Keeper (Oath-Keeper)}]
\emph{The word not spoken is a shield. The secret kept is a kindness. I will carry your truth so you do not have to.}

\vspace{2mm}
\noindent\textbf{Prerequisites:} Spirit 3+, Lore 2+, Rite of the Gentle Seal or Cantor of the Silent Choir.

\vspace{2mm}
\noindent\textbf{Confidant Techniques (choose as Talents):}

\begin{tblr}{
  colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.3,l]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Technique & Effect & Cost & Req. \\
\textbf{Calm Presence} & Once per scene, you may touch a creature suffering from Fear, Guilt, or Despair and remove that Condition. & 1 Fatigue & Spirit 3 \\
\textbf{Sealed Lips} & You may use the Hedge Gift \textit{Weight of the Unsaid} without spending XP (you know it instinctively). & Passive & Tier II \\
\textbf{Burden Steady} & You may carry one additional psychological Condition without stacking penalties. & Passive & Spirit 3 \\
\textbf{Confession’s Mercy} & When you use Rite of the Burden Bearer, the target’s relief lasts for the entire scene (instead of just the social roll bonus). & +1 Obligation & Tier III \\
\textbf{Sanctuary Keeper} & Once per session, you may declare a room you are in to be a \textit{Sanctuary} for one scene. All within gain +1 die to resist fear and coercion. & 1 Boon & Spirit 4, Tier III \\
\textbf{Final Vow} & Once per campaign, you may take a dying person’s last secret and carry it permanently. You gain +1 permanent Corruption, but you may never reveal that secret – not even under magical compulsion. & Free & Epic \\
\end{tblr}

\vspace{2mm}
\noindent\textbf{Master Technique (Epic Talent, 12 XP):}
\textit{The Unburdened Confessor.} Once per session, you may absorb a single traumatic memory or secret from a willing creature, erasing it from their mind permanently. You gain the memory as your own (with all emotional weight), and the creature is freed from its effects. You gain 1 permanent Corruption and mark +2 Obligation to the Silent Choir. You cannot use this talent again until you have spent a downtime scene in meditation or confession.
\end{tcolorbox}

\subsubsection*{Cantors \& Cults of the Silent Choir}

\textbf{The Cantor's Song.} A Cantor of the Silent Choir sings in \textit{breaths of omission} – melodies that are felt rather than heard, harmonies that resolve into silence. Their voice is used to quiet a room of gossip, to seal a confession with a hum that makes it sacred, or to place a gentle pressure on a tongue that would speak a harmful truth. The Choir's Cantors are often found in confessionals, in the rooms of the dying, and in the alcoves of great libraries, humming the songs that keep secrets safe.

\textbf{The Cult of the Sealed Lip.} The Silent Choir's cults meet in places where secrets are naturally kept: locked scriptoria, the back rooms of notaries, and the quiet corners of courthouses. The Sealed Lip is a fellowship of confessors, archivists, therapists, and spies who have learned that the most powerful word is the one left unspoken. A ritual of the Cult involves each member writing down a truth they wish they could tell, then sealing it in an envelope that is never opened. The envelopes are stored in a single locked chest – the \textit{Coffer of Unspoken Things}. Once per season, they pour the ashes of a burned confession into a small copper bell; the bell never rings, but they claim it “remembers” every secret fed to it. Their creed is written on the inside of the lid: \textit{"The word not spoken is a shield. The secret kept is a kindness."} Outsiders are welcome to bring a sealed envelope of their own, but they must never ask what is inside another’s.

\subsubsection*{Witchcraft of the Silent Choir (Lay / Secrecy)}

A witch who walks with the Silent Choir is a \textit{seal‑scribe} or \textit{lock‑keeper} – one who crafts the physical vessels of silence: a book whose pages refuse to reveal their contents to the wrong reader, a chest that cannot be opened without the keeper's whisper, or a letter that dissolves if read aloud. Their magic works through the materials of discretion: a quill that writes in ink that fades after one reading, a wax seal that screams if broken by an unauthorised hand, or a lock that only clicks shut when a secret is placed inside. Their tools are not hammers; they use candles, seals, and the soft pressure of a thumb on a clasp.

\textbf{The Witch’s Tool: The Candle of Unspoken Trust.} A thick, beeswax candle, grey as ash, that burns with a silent, smokeless flame. Once per session, the witch may light the candle and hold a sealed envelope, a door, or a person’s hand over the flame (not touching it). For the next hour, any secret placed in the envelope, spoken behind the door, or entrusted to the person cannot be revealed by magical means (compulsion, truth spells, divination). The candle gutters if the secret is about to be forced out, giving the keeper one minute to intervene. The candle melts entirely after nine uses.

\textbf{A Hedge Gift: Weight of the Unsaid (2 XP).} Once per scene, the witch may touch a person's lips and speak a single word (“hush”). For the next minute, that person cannot speak any word that would reveal a secret they have sworn to keep (including secrets they have overheard). If they try, they feel a mild pressure on their throat – enough to pause them, but not enough to cause harm. The effect ends if the witch speaks again.

\subsubsection*{The Silent Choir's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1.5cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Blessing & Burden / Quirk \\
1 & \textbf{Discretion's Gift:} +1 die to \textit{Notice} when someone is about to speak what should be kept. & \textbf{Silent Compulsion:} Urge to still tongues; --1 die to ordinary conversation. \\
2 & \textbf{Confidence Keeper:} Once/scene, shelter a secret from accidental spill. & \textbf{Secret Burden:} In places thick with confidences, suffer \textbf{Fatigue 1}. \\
3 & \textbf{Merciful Blindness:} Once/session, prevent discovery of a truth that would harm if revealed. & \textbf{Truth Sensitivity:} Lies grate; --1 die while surrounded by deception. \\
4 & \textbf{Sanctuary Sense:} Once/scene, veil a brief zone from eavesdropping. & \textbf{Communication Strain:} Others balk at full candour; --1 die to personal disclosure. \\
5 & \textbf{Wisdom of Restraint:} Once/session, see the full cost of airing any one truth. & \textbf{Silent Isolation:} --2 dice to form close bonds; you hold too much back. \\
6+ & \textbf{Choir's Voice:} Once/session, speak with the gathered prudence of unspoken truths. & \textbf{Truth Prison:} Mark +3 Obligation; risk becoming a living vault unable to share without harm. You cannot speak any secret you have carried unless you first mark 1 Fatigue to “loosen” it. \\
\end{tblr}

\subsection*{Playstyle Notes}
The Silent Choir favours guardians and confidants: those who stand between fragile lives and ruinous truth. Its rites fortify discretion, temper tempers, and build rooms where trust can breathe. The cost is weight: secrets you keep will keep you. Unlike the Witness (who would speak truth at any cost) or Ikasha (who hides for advantage), the Choir teaches that \textit{mercy is sometimes silence}. Ideal for therapists, confessors, archivists, and anyone who has ever chosen to carry a burden so that another might walk free.

\noindent\textbf{Emphasises}
\begin{itemize}
  \item \textbf{Protective Restraint:} silence used as shield
  \item \textbf{Confidential Guardianship:} duty to those who trust you
  \item \textbf{Merciful Omission:} letting some truths sleep
  \item \textbf{Discretion Wisdom:} judging what, when, and how to speak
  \item \textbf{Burden Bearing:} carrying heat so others do not burn
\end{itemize}

% --- Monastic Tradition: Whisper Monk (Silent Choir) ---
\begin{tcolorbox}[colback=black!5,colframe=gray!50!white,title=\textbf{Whisper Monk (Silent Choir)}]
  \emph{The word not spoken is a shield. The secret kept is a kindness.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Spirit 3+, Presence 2+, Rite of the Gentle Seal or Cantor of Silent Choir.
  
  \vspace{2mm}
  \noindent\textbf{Debt–Resistant Frame.} What is not said cannot be owed. Whisper monks gain \textbf{+1 die} to resist any magical debt or obligation that would force them to speak a secret, reveal a truth, or betray a confidence. If they succeed, the creditor hears only silence – and silence has no signature.
  
  \vspace{2mm}
  \noindent\textbf{Techniques:}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.4,l]},
    rowhead = 1, row{odd}={bg=gray!10}, row{1}={bg=gray!30}, hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Silencing Palm} & On an unarmed hit, the target cannot speak any word that would reveal a secret they have sworn to keep for one scene (Resolve DV 3 to resist). & 1 Fatigue & Spirit 3 \\
  \textbf{Calm Presence} & Once per scene, remove a Condition (Fear, Guilt, Despair) from a willing creature by touch. & 1 Boon & Spirit 3 \\
  \textbf{Gentle Seal} & Hush a space with respectful quiet. All present gain +1 die to actions of discretion, confidentiality, or de‑escalation this scene. & 1 Fatigue & Presence 2 \\
  \textbf{Burden Bearer} & Take another’s social or emotional weight for a time. Target gains +2 dice to social rolls while you suffer –1 die. & 1 Boon & Spirit 2 \\
  \textbf{Sanctuary Keeper} & Once per session, declare a room you are in to be a Sanctuary for one scene. All within gain +1 die to resist fear and coercion. & Free & Tier II \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{The Unburdened Confessor.} Once per session, absorb a single traumatic memory or secret from a willing creature, erasing it from their mind permanently. You gain the memory as your own (with all emotional weight), and the creature is freed from its effects. You may also ignore one debt or obligation that was based on that erased memory. Gain a permanent **Debt Scar** (a grey thread tied around your wrist that never loosens). The Hollow will no longer pursue that debt – but the memory may one day haunt you.
\end{tcolorbox}