% --- Patron: Thrysos, King of Revels (Ecstasy, Excess \& Salacious Delight) ---
\subsection{Thrysos, King of Revels (Ecstasy, Excess \& Salacious Delight)}

\textit{Lore.} Where torches flare and cups spill, where dancers stamp until dawn, there strides Thrysos crowned with ivy and horn. He is both host and beast, a patron of license who laughs at law. In his train follow the masked and the wild, the broken who dance to forget, the bold who revel to remember, and the lustful who seek the heat of bodies in the dark. His rites intoxicate, inflame, and unmask --- yet each cup leaves a mark, each kiss a chain, each night of abandon a dawn of reckoning.

Thrysos rules the spectrum from the grinning drunkard who sings bawdy songs in taverns to the ecstatic cults whose rites strip away not only clothes but every pretense of civilization. He is the god of the Bacchanalia, the wild hunt of the senses, the moment when the watcher becomes the participant and the participant becomes the sacrifice. In his milder aspect, he blesses the fertility of the harvest and the laughter of lovers. In his darker aspect, he drives maenads to tear apart the hero who refused to join the dance.

Among the peoples of Fate's Edge, Thrysos is called by many names. In Linn, his followers are the \textit{Storm-Dancers}, who whirl through gales until exhaustion, their cries mistaken for ecstasy. In Aelinnel, he is the \textit{Unbounded Guest}, whose arrival at a feast is marked by laughter, shattered glass, and at least one scandalous elopement. In Zakov, the \textit{Revel-King} is invoked before every pirate carouse, with the understanding that a hungover crew is a vulnerable one --- and that the captain's bed is never empty. In Vhasian hedge-courts, his name is whispered by those who seek a potion to lower inhibitions or a curse of unquenchable desire. His rituals are never secret --- they are too loud, too wet, too fragrant with spilled wine and sweat --- but they are always dangerous, for Thrysos demands that every excess be paid for, if not tonight then at the first hangover of spring, or in the unwanted child, or in the blood feud started by a jealous spouse.

\begin{quote}
\emph{``Drink, and be unafraid. Kiss, and be unbound. Dance, and be devoured. Every mask is true when worn long enough, and every truth is a mask for hunger.''}
\end{quote}

\noindent\textbf{Signs.} Wine that never sours; a mask that smiles without a wearer; goat-tracks circling thrice before vanishing; the sudden, inexplicable warmth of a stranger's hand on your arm; the sound of a zither played in an empty room; the scent of myrrh and sweat clinging to a garment that has never been worn.

\noindent\textbf{Cantor's Temptation.} Thrysos promises freedom without end: no oaths, no chains, only song and cup and flesh. Yet those who drink too deeply wake to find they serve not joy but appetite --- and appetite is a master that never sleeps.

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Ecstatic Release:} dance, intoxication, the shedding of inhibition (and clothing)
  \item \textbf{Masked Freedom:} identity as performance, truth through disguise, the thrill of the anonymous encounter
  \item \textbf{Feast and Famine:} excess, appetite, lust, and the cost of indulgence
  \item \textbf{Ritual Frenzy:} communal ecstasy that transforms participants into something feral and beautiful
  \item \textbf{The Bacchic Dark:} the point where pleasure tips into violence, where the maenad tears apart the intruder
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Thrysos does not bind a familiar of quiet habits. Their \textbf{Thiasos} is a small beast or mythological being of primal appetites: a satyr-kid with too-bright eyes that nibbles at wine-soaked bread, a fox that dances on its hind legs when the moon is full, a ferret that steals silk ribbons from lovers' bedposts, a gecko that clings to wine jugs and changes color with desire, or a small bronze cymbal that claps of its own accord during moments of flirtation. This creature requires no ordinary food, but it must be present during an act of indulgence (drinking, dancing, lovemaking, or feasting) each day. If neglected for a week, it becomes \textit{Compromised} (it shrieks at inappropriate moments, or its antics become grotesque, imposing --1 die on all social rolls involving subtlety or discretion for its bearer).

The \textbf{Codex} of Thrysos is never a sober book. It may be a wine-stained journal of erotic poetry, a set of ivory tablets carved with scenes of revelry, a leather folio containing pressed flowers from orgies, or a scroll that unravels further each time the reader laughs. Because Thrysos's truths are written in the body's language, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week adding a new verse, sketch, or memory of indulgence, or the Rites stored there become dull (treat as Compromised until reinvigorated with a night of celebration).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Thrysos may touch a held item (weapon, tool, or garment) and whisper a word of desire. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Performance} or \textit{Sway} rolls involving seduction, charm, or intimidation, and the first time each scene the bearer would be affected by a Fear effect, they may laugh it off (ignore it). After the scene, the user marks \textbf{+1 Obligation} as the craving for another taste of freedom gnaws at them.

\subsubsection*{Rites of Thrysos}

\paragraph{Rite of the Overflowing Cup (Low, 4 XP)} \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[SOCIAL]}, \texttt{[FEAR]}, \texttt{[AREA]} \\
\textbf{Materials:} A brimming vessel of wine (or a more intimate offering). \\
\textbf{Effect:} Allies who drink gain \textbf{+1 die} to social rolls this scene; enemies who drink suffer \textbf{--1 die}. Those who drink deeply may find their inhibitions loosened in other ways (GM's discretion for romantic or scandalous complications). \\
\textbf{Push It:} All in Near must test Resolve (DV 3) or be swept into revelry, possibly leading to unplanned liaisons or public confessions; mark \textbf{1 SB (Hearts)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Horned Dance (Low, 5 XP)} \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[BOOST]}, \texttt{[FOCUS]}, \texttt{[TRANSFORM]} \\
\textbf{Materials:} A mask or antlers (carved from a stag that died in rut). \\
\textbf{Effect:} Enter an ecstatic frenzy: \textbf{+2 dice} to melee or performance; ignore 1 Fatigue per round. Your movements become both seductive and threatening; those watching may be captivated or disturbed. \\
\textbf{Push It:} The frenzy spreads: one ally joins you in the dance, but both mark \textbf{1 SB (Spades)} as the lines between combat and carnality blur. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Breaking Mask (Standard, 7 XP)} \hfill \\
\emph{Instant; Near; No.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[DEBUFF]}, \texttt{[FEAR]} \\
\textbf{Materials:} A mask smashed in hand (or a lover's token torn in two). \\
\textbf{Effect:} Reveal the true face beneath the social mask: the target suffers \textbf{--2 dice} to disguise, deceit, or composure until scene end. They may also be compelled to admit a hidden desire (GM's discretion). \\
\textbf{Push It:} Truth is forced: the target must declare one hidden desire aloud, possibly to the wrong person; mark \textbf{1 SB (Diamonds)}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Wild Procession (Standard, 9 XP)} \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[SUMMON]}, \texttt{[WARD]}, \texttt{[AREA]} \\
\textbf{Materials:} A horn sounded thrice (or a kiss passed around a circle). \\
\textbf{Effect:} Summon spectral revelers --- satyrs, dancers, horned riders, and a few figures whose embraces are more than friendly. For the scene, the zone is \texttt{[WARD]}ed: hostile entry forces \textbf{1 SB (Spades)}; allies gain \textbf{+1 die} to resist compulsion or fear. The revelers may also grant a boon of Vigor to one couple (restore 1 Fatigue to two allies who share a bond of intimacy). \\
\textbf{Push It:} The procession spills beyond the zone; outsiders in Near must test Resolve (DV 4) or join the dance --- and perhaps the revel's other activities. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Endless Feast (High, 14 XP)} \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[BOOST]}, \texttt{[AREA]}, \texttt{[BURDEN]} \\
\textbf{Materials:} A feast laid for twelve, with wine, music, and comfortable cushions. \\
\textbf{Effect:} For one scene, the zone floods with plenty: allies restore Fatigue; all rolls tied to appetite or indulgence gain \textbf{+2 dice}. Leaving requires Resolve (DV 3) --- the comforts are hard to abandon. The feast also grants a temporary \textit{Heightened Desire} condition: the first seduction or intimidation roll each ally makes gains +1 die. \\
\textbf{Push It:} The feast persists as a hunger-haunt; start \textbf{Hedonist's Curse [6]}. Participants may find themselves unable to derive pleasure from mundane means, always craving the excess of that night. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 8 segments.

\paragraph{Rite of the Reveler's Apotheosis (High, 15 XP)} \hfill \\
\emph{Extended; Touch; No.} \\
\textbf{Tags:} \texttt{[TRANSFORM]}, \texttt{[BOOST]}, \texttt{[FATE]}, \texttt{[SACRIFICE]} \\
\textbf{Materials:} A crown of twisted vines that bleed when touched, the first laugh of a newborn, a promise broken at a crossroads, and the willing touch of a lover. \\
\textbf{Effect:} Grant the ultimate reveler's gift --- transformation into pure ecstatic essence. The target becomes immune to fear and despair, gains \textbf{+3 dice} to performance or frenzy, and can inspire allies within Near (\textbf{+1 die} to all rolls). Their aura also carries a scent of desire; anyone within Near who sees them must test Resolve (DV 3) or be mildly infatuated (disadvantage on hostile actions). \\
\textbf{Push It:} The transformation becomes contagious --- nearby allies may also transform but risk losing their humanity (their sense of self blurs into the ecstatic mob); mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\emph{Obligation:} 8 segments.

% --- Drunken Master Monk (Thrysos) ---
\begin{tcolorbox}[colback=black!5,colframe=green!50!black,title=\textbf{Drunken Master Monk (Thrysos)}]
  \emph{The true disciple does not fight the wine --- they let the wine fight for them. Every stagger is a feint, every spill a trip, every belch a battle cry, every wandering hand a distraction.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Body 3+, Spirit 3+, Rite of the Horned Dance learned or Cantor of Thrysos.
  
  \vspace{2mm}
  \noindent\textbf{Monastic Techniques (purchased as Talents):}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.1,l] X[1.3,l]},
    rowhead = 1,
    row{odd} = {bg=gray!10},
    row{1} = {bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Stagger Step} (Basic, 4 XP) & When an enemy misses you with a melee attack, you may immediately move one band (Close $\rightarrow$ Near or vice versa) without provoking opportunity attacks. & 1 Fatigue & Body 3 \\
  \textbf{Spill the Cup} (Basic, 5 XP) & As a reaction to being grappled or restrained, you may slip free and impose --1 die on the attacker's next roll. & 1 Boon & Dex 2 \\
  \textbf{Drunken Fist} (Basic, 5 XP) & Your unarmed strikes gain +1 die and ignore the first point of Armor when you are at Fatigue 2 or more (the wine ``guides'' you). & Passive & Body 3, Spirit 2 \\
  \textbf{Hangover Resilience} (Basic, 4 XP) & Once per scene, after suffering a complication or Story Beat from intoxication, you may clear 1 Fatigue (the headache fades, but the lesson remains). & Free & Spirit 3 \\
  \textbf{Reveler's Recovery} (Standard, 8 XP) & When you use Rite of the Horned Dance, you may also heal 1 Harm (not Condition) as the ecstasy knits flesh. & +1 Obligation & Tier II, Rite of the Horned Dance \\
  \textbf{Satyr's Sprawl} (Standard, 8 XP) & You are immune to being knocked prone while moving, and when you fall prone intentionally you may immediately stand as a free action. & Passive & Body 4, Athletics 2 \\
  \textbf{Bacchic Counter} (Standard, 9 XP) & When you succeed on a defense roll while at Desperate Position, you may immediately make a free unarmed attack against the attacker. & Once per scene & Tier III, Drunken Fist \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{The Unbroken Bottle.} Once per session, you may ignore all penalties from Fatigue for one scene. While this is active, each time you would be hit by a melee attack, you may instead stagger out of the way and force the attacker to hit another adjacent creature of your choice (narrate the stumble). After the scene ends, you collapse with Harm 2 (Exhaustion) and cannot use this talent again until you have rested for a full night.
\end{tcolorbox}
  
\subsubsection*{Cantors \& Cults of Thrysos}

\textbf{The Cantor's Song.} A Cantor of Thrysos sings in \textit{off-key harmonies} --- melodies that start wild and grow wilder, rhythms that mimic a galloping heart. Their voice is used to whip a crowd into a frenzy, to calm a hangover with a lullaby, or to make a mask cling to a face until the wearer is ready to take it off. Thrysos's Cantors are often found at festivals, in taverns at closing time, leading the procession of the horned dancers, or performing in private salons where the line between audience and participant is deliberately blurred.

\textbf{The Cult of the Horned Circle.} Thrysos's cults meet in clearings, barns, and the back rooms of inns after midnight. The Horned Circle is a fellowship of mask-carvers, vintners, musicians, courtesans, and those who have tasted the edge of ecstasy --- sometimes all in the same person. Their most sacred vessel is a hollowed-out log that was struck by lightning; they ferment wine inside it, and the wine \textit{whispers secrets} to those who drink it --- forgotten grudges, hidden desires, the true names of those who will betray them, and the faces of those who are willing. A ritual of the Circle involves each member donning a mask (hand-carved, never bought) and dancing until the first dancer falls from exhaustion. The fallen is crowned with ivy and declared ``guest of honor'' --- they lead the next circle, and sometimes the next bed. Their creed is carved into a single wine barrel that is never emptied: \textit{"The cup that never empties is the cup you share."} Outsiders are welcome to dance, but they must leave their mask behind when they go --- and if they drink from the lightning-log, they will hear the whispers for a year and a day, and they will return.

\textbf{The Maenads of the Crimson Caves (Bacchic Dark).} The darkest cult of Thrysos, the Maenads (or \textit{Bassarids}) are almost exclusively female, though men may join at the risk of becoming their prey. They gather in hidden caves or moonlit groves to perform the \textit{oreibasia} --- the mountain dance --- wearing fawnskins and carrying thyrsi tipped with bronze. Their rites begin with wine and music, escalate to ecstatic abandon, and sometimes end in \textit{sparagmos} --- the tearing apart of a living sacrifice (often a bull, but in the oldest stories, a human who refused to join the dance). This cult is banned in every civilized nation, but it persists in the wildest frontiers. A character who joins the Maenads gains access to darker rites, but also the risk of losing control entirely (the GM may treat this as a Corruption track that leads to a violent, permanent transformation).

\subsubsection*{Witchcraft of Thrysos (Lay / Revelry)}

A witch who walks with Thrysos is a \textit{mask-carver} or \textit{wine-smith} --- one who crafts the tools of ecstatic release: a mask that lets the wearer dance without tiring, a drum that beats in time with the drinker's pulse, a cup that never runs dry (but leaves a headache at dawn), or a perfume that makes the wearer irresistible for a single night. Their magic works through the materials of celebration: a thyrsus that marks the boundary of a sacred dance, a press that squeezes wine that ferments in hours instead of weeks, a needle that sews a costume that makes the wearer faster, louder, bolder, or a small clay lamp that burns with a flame that kindles desire. In Aeler, a True-Mason might carve a ``revel keystone'' into a feast hall, but a hedge-witch of Thrysos might instead drive a thyrsus into the floorboards to consecrate the space for an orgy.

\textbf{The Witch's Tool: The Thyrsus of the Unbound.} A staff of fennel or oak, topped with a pine cone and bound with ivy, often with a ribbon from a lover's garment. Once per session, the witch may strike the thyrsus against the ground three times. For the next scene, any creature within Near who is engaged in revelry (dancing, drinking, performing, or any act of intimacy) gains \textbf{+1 die} to social and physical actions, but any attempt to leave the zone requires a Resolve test (DV 3) --- the revel holds them. The thyrsus can also be used to draw a circle on the ground; any mask placed within the circle becomes ``true'' for one night --- the wearer cannot be recognized by anyone who was not present at the drawing, and their inhibitions are lowered (they may act on desires they would normally suppress).

\textbf{A Hedge Gift: Dancer's Cramp (Basic, 4 XP).} Once per scene, the witch may snap their fingers and whisper a dancer's rhythm. One target within Near range feels an irresistible urge to move --- they must tap a foot, sway, or take a step. This cannot force them into danger, but it can ruin stealth, break a staredown, or distract a guard for one exchange. The target may resist with Resolve (DV 3) to ignore the urge. A more salacious version (\textit{Lover's Cramp}) might be used in seduction scenes, forcing an unintended compliment or touch.

\subsubsection*{Thrysos's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1.5cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Gift & Corruption \\
1 & \textbf{Reveler's Grace:} +1 die to social rolls in feasts or crowds. & \textbf{Compulsive Toast:} Must drink or mark 1 Fatigue when drink is offered. \\
2 & \textbf{Horned Vitality:} Ignore the first 1 Fatigue each scene. & \textbf{Restless Appetite:} --1 die when abstaining from indulgence (food, drink, or touch). \\
3 & \textbf{Revel-Kin:} Once/scene, summon a minor revel-spirit (advantage die) --- a satyr, a maenad, or a muse of seduction. & \textbf{Masked Self:} Cannot act without a role or guise; --1 die to plain sincerity. \\
4 & \textbf{Bacchic Fury:} Once/session, frenzy grants +2 dice to melee for one exchange. Also, once per scene, you may ignore the effects of a single social faux pas (the revel covers it). & \textbf{Ecstatic Blindness:} --1 die to perception outside the revel; you may not notice threats that require sobriety. \\
5 & \textbf{Feast Without End:} Once/session, allies in Near clear all Fatigue. & \textbf{Hungering Host:} You consume twice as much; denial causes Harm 1 (Stress). You may also develop a reputation: every feast you attend ends in scandal. \\
6+ & \textbf{Satyr King's Mantle:} For one scene, become an avatar of revels; all allies gain +2 dice, enemies suffer --2 dice. Your presence also inspires desire: once during the scene, you may compel one target to act on a hidden passion (Resolve DV 4). & \textbf{Tyrant of Pleasure:} Mark +3 Obligation; the revel grows monstrous, demanding a sacrifice (an oath, a relationship, or a life). \\
& & \textbf{Beast of Endless Revel:} Permanent transformation --- your mortal form becomes optional; must make Presence+Performance (DV 4) to appear civilized or risk terrifying (or entrancing) all who see you into joining your eternal celebration. You are always at least mildly intoxicated, and sobriety causes you physical pain (1 Fatigue per hour). \\
\end{tblr}

\begin{quote}
``Let the wine flow, let the masks fall, let the fire burn away all restraint. In the revel we find truth --- raw, wild, and beautiful. And in the aftermath, we find the hunger for more.''
\end{quote}

\subsection*{Playstyle Notes}
Thrysos rewards those who embrace risk, abandon convention, and find power in letting go --- both the laughter of the drunkard and the ecstasy of the maenad. His followers become masters of ecstatic ritual, social manipulation through celebration and seduction, and the dangerous art of losing control. The corruption progression leads toward a beast of endless revel --- powerful, charismatic, but never truly sober, never truly at rest, and always hungry. Ideal for players who enjoy wild scenes, social chaos, characters who dance on the edge of self-destruction, and those who wish to explore the full spectrum of hedonism from playful flirtation to violent frenzy.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Ecstatic Release:} music, dance, intoxication, the shedding of inhibition (and clothing)
  \item \textbf{Masked Freedom:} identity as performance, the thrill of the anonymous encounter
  \item \textbf{Feast and Famine:} excess, appetite, lust, and the cost of indulgence
  \item \textbf{Ritual Frenzy:} communal ecstasy that transforms participants into something feral and beautiful
  \item \textbf{The Bacchic Dark:} the point where pleasure tips into violence, where the revel demands blood
  \item \textbf{Seduction as Weapon:} using desire to manipulate, distract, or destroy
\end{itemize}

% --- Monastic Tradition: Drunken Master (Thrysos) ---
\begin{tcolorbox}[colback=black!5,colframe=green!50!black,title=\textbf{Drunken Master (Thrysos)}]
  \emph{The true disciple does not fight the wine --- they let the wine fight for them. Every stagger is a feint, every spill a trip, every belch a battle cry.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Body 3+, Spirit 3+, Rite of the Horned Dance or Cantor of Thrysos.
  
  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} Thrysos's monks owe nothing but the next glass. They gain \textbf{+1 die} to ignore or resist any magical debt, geas, or contract that would compel them to act against their revelry. This boon can be used once per scene, but each use leaves a faint scent of wine on the monk --- the Hollow may find them easier to track (GM gains 1 SB).
  
  \vspace{2mm}
  \noindent\textbf{Techniques (purchased as Talents):}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.1,l] X[1.3,l]},
    rowhead = 1,
    row{odd} = {bg=gray!10},
    row{1} = {bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Stagger Step} (Basic, 4 XP) & When an enemy misses you with a melee attack, move one band (Close$\leftrightarrow$Near) without provoking opportunity attacks. & 1 Fatigue & Body 3 \\
  \textbf{Spill the Cup} (Basic, 5 XP) & As a reaction to being grappled, slip free and impose --1 die on the attacker's next roll. & 1 Boon & Dex 2 \\
  \textbf{Drunken Fist} (Basic, 5 XP) & Unarmed strikes gain +1 die and ignore 1 Armor when you are at Fatigue 2+. & Passive & Body 3, Spirit 2 \\
  \textbf{Hangover Resilience} (Basic, 4 XP) & Once per scene, after a complication from intoxication, clear 1 Fatigue. & Free & Spirit 3 \\
  \textbf{Satyr's Sprawl} (Standard, 8 XP) & Immune to being knocked prone while moving; standing from prone is a free action. & Passive & Body 4, Athletics 2 \\
  \textbf{Bacchic Counter} (Standard, 9 XP) & When you succeed on a Desperate defense roll, make a free unarmed attack against the attacker. & Once per scene & Tier III, Drunken Fist \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{The Unbroken Bottle.} Once per session, ignore all Fatigue penalties for one scene. When you would be hit by a melee attack, you may force the attacker to hit another adjacent creature of your choice. After the scene, you suffer Harm 2 (Exhaustion). You gain a permanent \textbf{Debt Scar} --- a purple bruise shaped like a grape cluster --- but may ignore one forced Obligation or debt repayment that targets you (the patron's mark confuses the creditor).
\end{tcolorbox}