\subsection{Umande, Dew-Mother of the Living Canopy (Growth \& Shelter)}
\begin{quote}\itshape
“Walk softly and you will hear the roots speaking of where to stand, when to bend, and whom to shelter.”
\end{quote}

\paragraph{Lore.}
Across the humid belts from the baobab-studded savannas to banyan-latticed monsoon forests, travelers whisper of \textbf{Umande}, the breath on the leaves at dawn. She is the pact of shade and water: a patient patron who measures worth in what you protect and what you prune. Her ministers are the unseen—lichen, mycelium, creepers—knitting harm back into wholeness, yet strangling cruelty where it clings too long. Umande favors wardens, foragers, midwives, and wayfarers who leave places better than they found them.

\paragraph{Patron’s Gift — Verdant Bond.}
Your sworn implement (staff, spear, billhook, kukri, or similar) knots itself in living vine and hard sap.
\begin{itemize}
  \item \textbf{Enchanted Weapon:} Your sworn implement counts as an enchanted melee weapon \textbf{+1 Melee}.
  \item \textbf{Thematic Skill:} You gain \textbf{+1 to Survival} while carrying any living sprig, seed, or leaf from your sanctum or a protected grove.
\end{itemize}

% =========================
% RITES (follow Ikasha style)
% =========================

\paragraph{Dawn’s Condensation (Low, 4 XP)} \emph{Scene; Touch/Self; Yes (resist only).}\\
\textbf{Materials:} A leaf cupping a few drops of clean water (or a wetted cloth).\\
\textbf{Effect:} Wick dew into flesh. Remove one minor Condition related to fatigue, heat, or dehydration \emph{or} grant +1 die to the next Survival or Medicine test this scene.\\
\textbf{Push It:} Also clear 1 SB of environmental stress (thirst/heat) for one ally within reach.\\
\textbf{Backlash (Life/Earth):} Skin chills and prunes; you suffer --1 die to Presence tests until you warm up.

\paragraph{Root-Weave Aegis (Low, 4 XP)} \emph{Scene; Near (close range); No.}\\
\textbf{Materials:} A twist of fiber, grass cord, or mycelial thread.\\
\textbf{Effect:} Sprout a knee-high barricade of roots and woven stems in a 2m arc. Gain \textbf{Cover} against ranged attacks and upgrade Position by one step for holds/defense while adjacent.\\
\textbf{Push It:} The first enemy crossing the weave is \textbf{Hindered} (movement penalty) this exchange.\\
\textbf{Backlash (Earth/Water):} Soil turns slick; you suffer --1 die on your next movement check this scene.

\paragraph{Whispering Mycelium (Standard, 8 XP)} \emph{Scene; Far (site-scale); Yes.}\\
\textbf{Materials:} A coin-sized fungus cap or a pinch of sporeprint.\\
\textbf{Effect:} Lay fingertips to soil or wood and commune with the hyphal web. Ask \textbf{two} questions about recent passage, injury, or decay within a circa 100m radius (GM answers truthfully but from the web’s perspective). Gain \textbf{advantage} on one track, forage, or concealment action tied to that intel this scene.\\
\textbf{Push It:} Ask a third question, or extend to 300m.\\
\textbf{Backlash (Life):} Your senses blur with rot; --1 die to non-Survival observation until you take a breath and steady yourself.

\paragraph{Green Path Unfurling (Standard, 8 XP)} \emph{Scene; Line of sight; Yes.}\\
\textbf{Materials:} A fresh tendril or vine.\\
\textbf{Effect:} Vegetation parts and cushions your passage. For the scene, you and up to \textbf{two} allies ignore minor difficult terrain in natural growth and gain +1 die to Stealth \emph{or} Athletics when moving through foliage.\\
\textbf{Push It:} Also \textbf{confound pursuit}: the first attempt to track you through this area is at --1 die.\\
\textbf{Backlash (Earth):} You leave telltale chlorophyll stains; tracking you in open ground gains +1 die until the next rest.

\paragraph{Rain-Caller’s Oath (Major, 12 XP)} \emph{Scene; Region (sky above); Yes.}\\
\textbf{Materials:} A hollow seedpod with three droplets of clean water inside.\\
\textbf{Effect:} Petition Umande to gather clouds or break them. Choose one:
\begin{itemize}
  \item \textbf{Summon Gentle Rains:} Over the next hour, steady rain falls; fires weaken, dust settles, and heat penalties are suppressed. Gain +1 die to Survival/Medicine to aid many.
  \item \textbf{Part the Canopy of Storms:} Winds ease and rainfall slackens to a workable drizzle; ranged penalties from heavy weather lessen by one step.
\end{itemize}
\textbf{Push It:} Shape the rain’s \emph{band}: center it on a location you can name within sight, sparing adjacent fields or focusing on a blaze.\\
\textbf{Backlash (Water/Life):} The sky takes its price; you suffer a \textbf{Drained} minor Condition (until warm food/rest), and plants nearby leach a hint of color.

\paragraph{Banyan’s Embrace (Major, 12 XP)} \emph{Instant; Near; Yes.}\\
\textbf{Materials:} A looped root or braided cord.\\
\textbf{Effect:} Living cords erupt to \textbf{Grapple} a single foe or to \textbf{Brace} a structure (choose one):
\begin{itemize}
  \item \textbf{Grapple:} Target must beat DV 4 (Body or Melee) to break free; on failure, they are Restrained until end of next exchange.
  \item \textbf{Brace:} A listing cart, palisade, or bridge segment holds; immediate collapse/damage is postponed; gain time to act.
\end{itemize}
\textbf{Push It:} Affect up to \textbf{two} adjacent targets/segments.\\
\textbf{Backlash (Earth):} Your limbs ache like old wood; --1 die to Melee checks until you stretch or rest.

\paragraph{Sanctum of the Green Court (Epic, 16 XP)} \emph{Ritual; Hours; Circle; Yes.}\\
\textbf{Materials:} Four living stakes (saplings) planted on the quarter points; a bowl of spring water; a woven crown of grass.\\
\textbf{Effect:} Consecrate a \textbf{Grove-Sanctum} (small campsite or shrine). While within, allies gain +1 die to \textbf{Survival} and \textbf{Medicine}; hostile entities of rot, blight, or open flame suffer --1 die to actions that harm living plants. Once per session within the sanctum, you may \textbf{negate} one minor environmental Complication (smoke, thirst, exposure).\\
\textbf{Push It:} Bind a \textbf{Vow}: name a creature, kin-group, or landmark under the Grove’s protection; the first direct harm against that Vow inside the sanctum immediately generates 1 SB for the GM \emph{and} grants you 1 Boon.\\
\textbf{Backlash (Life/Earth):} The grove remembers. If you break the Vow, Umande withholds dew: you cannot Push Umande rites until you atone (meaningful restoration or protection scene).

% Optional flavor epithets for different regions
\paragraph{Epithets.} In the eastern monsoon lands, she is \textit{She-Who-Hangs-The-Bridges} among banyan roots; on the savannas, \textit{Matron of Baobabs}; in mist forests, \textit{Mother-of-Dew} who writes omens on leaves at dawn.
