% --- Patron: Varnek Karn, the Death's Negotiator (Bargains with the Departed) ---
\subsection{Varnek Karn, the Death's Negotiator (Bargains with the Departed)}

\textit{Lore.} Varnek Karn dwells in the grey between life and death, where the newly departed linger with unfinished business and elder shades guard secrets dearer than gold. He is the broker who speaks across the veil, the mediator of disputes between living and dead, and the archivist of every promise sworn beyond the grave. His servants practice post‑mortem diplomacy: they negotiate, consult, and contract with spirits as equals---mindful that the dead are patient and rarely generous.

Yet Varnek Karn is also the patron of those who *raise* the dead, not through force but through bargain. The dead may be persuaded to walk again, to fight for a season, or to guard a threshold, but every such compact is a debt. The dead demand their due: a memory, a future, a name, or simply the promise that the living will remember them. Those who serve Varnek Karn learn that raising a spirit is never a command; it is a negotiation, and the spirit always has a price.

Among the peoples of Fate's Edge, Varnek Karn is the patron of those who treat with the departed respectfully. In Aeler, he is the \textit{Breath‑Accountant}, who audits the spiritual ledgers of the Deep Vaults and ensures that the dead are not forgotten. In Ykrul, he is the \textit{Bone‑Speaker}, whose drum calls ancestors to council before any major migration. In Vhasia, he is the \textit{Silent Notary}, who witnesses death‑bed testaments and ensures their terms are honoured. In the Mistlands, he is the \textit{Bell‑Shrouded}, whose unknotted rope holds the fog at bay. Wherever the living and the dead must negotiate, Varnek Karn's servants are there – not to command, but to listen.

\begin{quote}
“The dead have all the time in the world to consider their offers, and no need to rush to agreement. They know what the living forgot, and remember every slight with perfect clarity. In Varnek’s halls, both parties pay a price—but the currency is never what you expect.”
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Post‑Mortem Diplomacy:} parleys with ghosts and recent dead
  \item \textbf{Ancestral Wisdom:} counsel from forebears and keeping of funerary law
  \item \textbf{Unfinished Business:} resolving the concerns that bind spirits
  \item \textbf{Death‑Bound Contracts:} obligations that outlive the flesh
  \item \textbf{The Unquiet Host:} raising temporary armies of the dead through pact
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Varnek Karn does not bind a living familiar. Their \textbf{Thiasos} is a \textit{remnant of the honoured dead}: a lead seal that has witnessed a funeral, a knucklebone from an ancestor who died content, or a small iron bell that rings only when a spirit is willing to speak. This object requires no food, but it must be placed on a grave or threshold at dusk each day. If neglected for a week, the Thiasos becomes \textit{Compromised} (it emits a faint keening that imposes −1 die on all social rolls with the living).

The \textbf{Codex} of Varnek Karn is never a living book. It may be a slate carved with names that fade and reappear, a bundle of burial tablets, or the inside of a lead‑lined casket. Because the dead are not bound by paper, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week recording a new death or a settled dispute, or the Rites stored there become illegible (treat as Compromised until restored).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Varnek Karn may touch a held item (weapon, tool, or funerary token) and whisper a death‑bargain. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Lore} or \textit{Investigation} rolls involving death, ancestry, or inheritance, and the bearer may ask one question of a corpse less than a day dead (it answers with a single word or a gesture). After the scene, the user marks \textbf{+1 Obligation} as the weight of the grave lingers.

\subsubsection*{Rites of Varnek Karn}

\paragraph{Rite of the Crossing Fee (Low, 4 XP)} \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[SPIRIT]}, \texttt{[TRUTH]} \\
\textbf{Materials:} A coin for the tongue of the deceased, or a lock of hair from the bereaved. \\
\textbf{Effect:} Open speech with a specific dead individual. Gain \textbf{+1 die} to \textit{Investigation} or \textit{Lore} when seeking knowledge they once held. \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Ancestral Consultation (Low, 5 XP)} \hfill \\
\emph{Scene; Self; Yes.} \\
\textbf{Tags:} \texttt{[SPIRIT]}, \texttt{[TRUTH]}, \texttt{[FATE]} \\
\textbf{Materials:} A personal token of the departed, or their favoured food/drink. \\
\textbf{Effect:} Ask one question of an ancestral spirit regarding family history, hidden legacies, or inherited debts. \\
\textbf{Push It:} The spirit grants detailed, prophetic guidance but exacts a small favour; mark \textbf{1 SB (Hearts)}. \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Mediated Pact (Standard, 8 XP)} \hfill \\
\emph{Scene; Near; Yes.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[BOND]}, \texttt{[COMMAND]}, \texttt{[SPIRIT]} \\
\textbf{Materials:} Two copies of a written accord (one buried with the dead, one kept by the living), or a circle of salt mixed with grave‑dust. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Death‑Bound Pact:} Bind a compact between a living signatory and a willing spirit. Both gain \textbf{+1 die} when fulfilling the terms. Breach angers the spirit and Varnek both.
  \item \textbf{Spirit Medium:} Serve as a conduit for several spirits to speak. Gain \textbf{+2 dice} to mediate disputes between living and dead and to facilitate group negotiation.
\end{itemize}
\textbf{Push It:} If you chose \textit{Pact}, the pact enforces itself: a breach triggers supernatural reprisal; mark \textbf{+1 Obligation}. If you chose \textit{Medium}, a potent spirit rides your voice to speak directly; mark \textbf{Exposure +1}. \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Unquiet Host (Standard, 10 XP)} \hfill \\
\emph{Scene; Zone; Yes.} \\
\textbf{Tags:} \texttt{[SUMMON]}, \texttt{[AREA]}, \texttt{[BOND]}, \texttt{[DEBT]} \\
\textbf{Materials:} A battlefield relic (a broken sword, a skull, a scrap of uniform) and a handful of grave‑earth. \\
\textbf{Effect:} You negotiate with the lingering dead of a place. Summon a \textbf{Temporary Host} – a spectral warband treated as a single \textbf{Cap 3 Follower} with the following traits:
\begin{itemize}
  \item \textbf{Assist:} May grant \textbf{+3 dice} to a single combat or intimidation roll, or provide a free \textit{Setup} action (e.g., “they distract the guards”).
  \item \textbf{Independent Action:} Once per scene, the Host may take one action from the Follower list (e.g., \textit{Scout \& Signal}, \textit{Distract \& Draw}) at the cost of marking \textbf{Exposure +1}.
  \item \textbf{Weakness:} The Host is vulnerable to salt, bells, and consecrated ground. Any attack using such materials ignores their spectral nature.
\end{itemize}
The Host lasts for the remainder of the current scene, or until dismissed. When it departs, the dead demand their due: you must either mark \textbf{+2 Obligation} or perform a \textit{Memorial Rite} before the next dawn (a downtime action, no roll). If you fail to do so, the Host returns unbidden at the worst moment, acting as a hostile hazard (GM’s discretion). \\
\textbf{Push It:} Extend the Host’s duration to the entire session. The debt is doubled: mark \textbf{+4 Obligation} or perform a memorial rite \textit{and} sacrifice a personal memory (erase a minor fact or a Bond for one session). \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Final Testament (High, 13 XP)} \hfill \\
\emph{Extended; Touch; Yes.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[FATE]}, \texttt{[COMMAND]}, \texttt{[DEATH]} \\
\textbf{Materials:} A dying person’s last words recorded in their blood, witnessed by you. \\
\textbf{Effect:} Seal the last request with weight beyond life. Targets must pass \textit{Resolve} (DV 4) to ignore it. \\
\textbf{Push It:} The testament becomes a geas: only fulfilment—or Varnek’s wrath—can break it; mark \textbf{+2 Obligation}. \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Eternal Archive (High, 14 XP)} \hfill \\
\emph{Extended; Zone; Yes.} \\
\textbf{Tags:} \texttt{[TRUTH]}, \texttt{[AREA]}, \texttt{[FATE]}, \texttt{[SANCTUARY]} \\
\textbf{Materials:} A library mortared with tomb‑stone, ink from cremation ash. \\
\textbf{Effect:} Found a repository where the dead deposit knowledge. Gain a permanent \textbf{+1 die} to \textit{Lore} for historical or genealogical research made within. Spirits periodically contribute testimonies. \\
\textbf{Push It:} The archive awakens and begins to demand curatorship and payment; mark \textbf{Harm 1 (Stress)}. \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 8 segments.

\subsubsection*{Necromantic Summoning (Varnek Karn’s Way)}

For summoners (Pact‑Whisperers) who follow Varnek Karn, the usual Leash mechanics gain an additional option. When you call a spirit (using the Pact‑Whisperer rules), you may choose to bind it with a \textbf{Death‑Bound Contract} instead of the standard Leash. Instead of ticking Leash on harm or disobedience, the spirit gains a \textbf{Promise Clock [4]} named “Debt Unpaid.” Each time it acts against your interests, you may advance the clock by 1. When the clock fills, the spirit must either obey one final command (even against its nature) or immediately depart and become an enemy. However, while under contract, the spirit cannot be dismissed early – you are committed. This replaces the Boon Finesse option; you cannot spend Boons to clear ticks. This method is slower but more reliable for extended servitude, and it fits Varnek Karn’s theme of negotiation over coercion.

\subsubsection*{Cantors \& Cults of Varnek Karn}

\textbf{The Cantor’s Song.} A Cantor of Varnek Karn sings in \textit{dirges of negotiation} – melodies that invite the dead to speak, harmonies that soothe restless shades. Their voice is used to call an ancestor to council, to lull a vengeful ghost into recounting its grievance, or to sing a funeral procession past a haunted crossroads. Varnek’s Cantors are often found in graveyards, in the halls of the dying, and at the tables where inheritance disputes are settled, their songs bridging the silence between breath and decay.

\textbf{The Cult of the Unforgotten Ledger.} Varnek Karn’s cults gather in crypts, ossuaries, and the quiet rooms above funeral parlours – but one cell, known as the \textit{Ink‑Flesh Communion}, tattoos the names of the dead onto their own skin, believing that each name grants a fragment of the ancestor’s memory. Their ritual involves writing a death‑record with a needle dipped in ash and wine; the tattoo whispers the dead person’s last words when pressed. The Unforgotten Ledger itself is a single, ever‑growing book bound in grey hide – it is said to \textit{whisper} at night, adding names without a scribe. Outsiders are welcome to bring a name of someone they have lost; they may read the ledger but not write in it. Their creed: \textit{“The dead are not gone. They have only changed their address. Be polite, pay your respects, and never make a promise you cannot keep.”}

\subsubsection*{Witchcraft of Varnek Karn (Lay / Funerary)}

A witch who walks with Varnek Karn is a \textit{funeral notary}, a \textit{grave‑warden}, or a \textit{memory‑keeper} – a lay practitioner who uses ink, seals, and small rites to witness the passage of the dead. Their magic is the art of recording and releasing: a leaden seal that certifies a death‑bed promise, a quill that writes in invisible ink that appears only when a ghost is near, or a carved wooden stamp that quiets a restless spirit for a night. In Aeler, a True‑Mason might carve a “memory niche” into a crypt wall, but a hedge‑witch of Varnek Karn might press a wax seal onto the niche’s edge, binding the dead to speak truth when the seal is broken. In Zakov, a net‑mender might weave a shroud with a single loose thread; pulling it releases the drowned sailor’s soul to the tide.

\textbf{The Witch’s Tool: The Seal of the Silent Witness.} A leaden seal on a leather thong, its face carved with a single closed eye. Once per session, the witch may press the seal into warm wax (or directly onto a stone threshold) while speaking the name of a dead person. For the next day, anyone who speaks that name near the seal hears a faint whisper – a single word or a sigh – from the deceased. The seal may also be used to “witness” a death‑bed promise: pressing it onto a written testament makes the document unbearably heavy to anyone who would break the vow.

\textbf{A Hedge Gift: Echo of the Honoured Dead (2 XP).} Once per scene, the witch may touch a physical remnant of the dead (a bone, a lock of hair, a piece of clothing) and whisper a question. For a moment, the remnant holds a memory – the GM answers one factual question about the dead person’s life or death, but the answer is always tinged with the emotion of the moment (e.g., “she was afraid,” “he was angry,” “they were at peace”).

\subsubsection*{Varnek Karn’s Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1.5cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Death’s Whisper:} +1 die to \textit{Notice} when the dead seek to communicate. & \textbf{Graveyard Pallor:} –1 die to social rolls with the living; your presence unsettles. \\
2 & \textbf{Ancestral Insight:} Once/scene, gain +2 dice for family history or inherited troubles. & \textbf{Spectral Awareness:} Ever‑aware of nearby spirits; suffer \textbf{Fatigue 1} in places of many deaths. \\
3 & \textbf{Medium’s Gift:} Once/session, act as a flawless conduit for a spirit. & \textbf{Death‑Touched:} Cold spots and small portents follow you; –1 die to ordinary social grace. \\
4 & \textbf{Final Authority:} Once/scene, compel adherence to a true last request (Resolve DV 4). & \textbf{Corpse‑Pale:} Visibly wan; –2 dice where warmth or vigour is required. \\
5 & \textbf{Archive Keeper:} Once/session, access perfect recall of a historical event or family secret. & \textbf{Unquiet Presence:} The dead trail you; –1 die to stealth and to surprising the living. \\
6+ & \textbf{Death’s Confidant:} Once/session, compel a willing spirit to answer three questions truthfully. & \textbf{Halfway House:} Mark +2 Obligation; risk becoming half‑withdrawn from the living world. You must perform a funeral rite for a stranger before the next dawn, or mark an additional +2 Obligation. \\
\end{tblr}

\subsection*{Playstyle Notes}
Varnek Karn favours investigators, arbiters, and heralds of funerary law. Power grows through cultural sensitivity, shrewd bargaining, and careful record‑keeping; the price is a thinning tie to warmth and mortal ease. Expect mysteries of lineage, obligations that outlive their makers, and parleys where both sides of the veil demand their due. The \textit{Rite of the Unquiet Host} allows a follower to temporarily command a small army of the dead, but the debt incurred ensures that such power is never free – and the dead always remember who called them.

\noindent\textbf{Varnek Karn emphasises}
\begin{itemize}
  \item \textbf{Diplomacy over Dominion:} bargaining with the dead rather than commanding them
  \item \textbf{Knowledge Brokerage:} trading for secrets and testimonies beyond the grave
  \item \textbf{Cultural Protocols:} rites, tokens, and proper doors for speaking with ancestors
  \item \textbf{Mediation:} resolving conflicts that bind the living to the dead
  \item \textbf{Historical Continuity:} showing how past vows shape the present
  \item \textbf{The Unquiet Host:} raising temporary armies through pact, not domination
\end{itemize}

% --- Monastic Tradition: Bone Monk (Varnek Karn) ---
\begin{tcolorbox}[colback=black!5,colframe=gray!70!black,title=\textbf{Bone Monk (Varnek Karn)}]
  \emph{The dead have all the time in the world to consider their offers.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Spirit 3+, Resolve 2+, Rite of the Crossing Fee or Cantor of Varnek Karn.
  
  \vspace{2mm}
  \noindent\textbf{Debt–Resistant Frame.} The dead do not collect – they witness. Bone monks gain \textbf{+1 die} to resist any magical debt or obligation that would bind them beyond the grave. If they succeed, the creditor feels a cold hand on their shoulder – a reminder that death cancels all ledgers.
  
  \vspace{2mm}
  \noindent\textbf{Techniques:}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.4,l]},
    rowhead = 1, row{odd}={bg=gray!10}, row{1}={bg=gray!30}, hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Ancestor's Knuckle} & Reroll a missed unarmed attack. The ancestor who guides your fist demands a name or a memory after the scene. & Free (once/scene) & Spirit 3 \\
  \textbf{Ancestral Consultation} & Ask one question of an ancestral spirit regarding family history, hidden legacies, or inherited debts. & 1 Fatigue & Lore 2 \\
  \textbf{Grace of the Grave} & When you would take a Condition from fear or despair, you may ignore it – a dead relative steadies your hand. & 1 Boon & Spirit 2 \\
  \textbf{Death's Whisper} & +1 die to Notice when the dead seek to communicate. You may also speak with a corpse less than a day dead (one question). & Passive & Spirit 3 \\
  \textbf{Final Testament} & Seal a dying person’s last request with weight beyond life. Targets must pass Resolve (DV 4) to ignore it. & 1 Fatigue & Tier II \\
  \textbf{Unquiet Host} & Negotiate with lingering dead of a place. Summon a spectral warband (Cap 3) for one scene. When it departs, you must mark 1 Fatigue or perform a Memorial Rite. & 1 Boon & Tier III \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Death's Confidant.} Once per session, compel a willing spirit to answer three questions truthfully. You may also ignore one debt that would have been owed to a deceased creditor – the ledger entry becomes dust. Gain a permanent **Debt Scar** (your fingertips are cold to the touch) and the Hollow notes your name, but you may call on one ancestor to resist a future obligation (the dead know how to wait).
\end{tcolorbox}