% --- Patron: Vorthak, the Hunger Unbound (Consumption \& Metamorphic Hunger) ---
\subsection{Vorthak, the Hunger Unbound (Consumption \& Metamorphic Hunger)}

\textit{Lore.} Vorthak is the principle of absolute consumption: an appetite that devours not only what is, but the very potential of what might be. Each feeding transforms both eater and eaten. His servants become conduits of metamorphosis through ingestion, learning to consume not only matter but also properties, identities, and even strands of fate. Power comes by taking in --- and the cost is becoming more like the hunger you channel.

Among the peoples of Fate's Edge, Vorthak is known by many fearful names. In Aeler, he is the \textit{Deep Forge}, the hunger of the mountain that consumes ore and returns steel. In Zakov, he is the \textit{Salt-Hunger}, the appetite that preserves by consuming the moisture from flesh. In Linn, he is the \textit{Tide-That-Eats-The-Shore}, a slow, patient consumption that reshapes coastlines. His followers are rarely beloved, but they are essential --- for without consumption, there is no transformation, and without transformation, there is only stagnation.

\begin{quote}
``To consume is to become. To devour is to transform. In Vorthak's embrace, the boundary between eater and eaten dissolves into something new --- and something hungry.''
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  \item \textbf{Metamorphic Consumption:} transformation through ingestion of matter, energy, or concepts
  \item \textbf{Essence Absorption:} gaining properties by consuming them utterly
  \item \textbf{Hunger Propagation:} spreading consumption as a force of change
  \item \textbf{Identity Metabolism:} consuming and adopting aspects of others' natures
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}
A Runekeeper sworn to Vorthak does not bind a familiar of flesh alone --- their \textbf{Thiasos} is a small creature of endless appetite: a glossy black beetle that devours rot and grows more lustrous with each meal, a pale grub that pupates into a moth only after consuming something precious, a leech that clings to the shadow of the wounded, a small eel that lives in the bilge and grows an eye each time its master consumes an essence, or a rat with a second mouth that whispers the names of things recently devoured. This creature requires no food in the ordinary sense, but it must be allowed to consume something \textit{significant} (a memory, a secret, a physical token) each week. If neglected, it becomes \textit{Compromised} (it gnaws at its bearer's belongings, consumes a stray thought, and imposes --1 die on Concentration or memory rolls).

The \textbf{Codex} of Vorthak is never a static record. It may be a stomach-bound journal written on cured hide, its pages growing thicker each time a Rite is performed; a set of teeth strung on a leather cord, each tooth etched with a consumed spell; or a wooden bowl stained with the residue of a thousand meals, the grain of the wood shifting as new essences are added. Because hunger is never satisfied, the Codex requires \textit{upkeep}: the Runekeeper must spend an hour each week feeding it --- a drop of blood, a burnt letter, a morsel of rare meat --- or the Rites stored there become faint (treat as Compromised until re-fed).

\subsubsection*{Patron's Gift (Imbuement)}
Once per scene as an action, a Runekeeper of Vorthak may touch a held item (weapon, tool, or piece of armor) and whisper a word of consumption. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Survival} or \textit{Craft} rolls involving organic or consumable materials, and the first time each scene the bearer would take a Condition from poison, disease, or starvation, they may ignore it (the hunger consumes the affliction). After the scene, the user marks \textbf{+1 Obligation} as the scent of turned earth and ash clings to them.

\subsubsection*{Rites of Vorthak}

\paragraph{Rite of the Hungry Touch (Low, 5 XP)} \hfill \\
\emph{Scene; Touch/Self; Yes.} \\
\textbf{Tags:} \texttt{[CONSUME]}, \texttt{[BOOST]}, \texttt{[SENSE]}, \texttt{[RISK]} \\
\textbf{Materials:} A fragment of the thing to be consumed (food, substance, symbol); or an empty vessel or a place where something was devoured. \\
\textbf{Effect:} Choose one:
\begin{itemize}
  \item \textbf{Metabolic Touch:} Absorb an essential trait of the target. Gain \textbf{+1 die} to actions aligned with the absorbed property for one scene.
  \item \textbf{Hunger's Whisper:} Sense what in the area may be consumed and the benefit it offers. Gain \textbf{+1 die} to \textit{Investigation} when seeking consumable resources.
\end{itemize}
\textbf{Push It:} (Whisper only) The hunger turns contagious: nearby feel your appetite; mark \textbf{1 SB (Clubs)}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Essence Feast (Standard, 8 XP)} \hfill \\
\emph{Scene; Touch; Yes.} \\
\textbf{Tags:} \texttt{[CONSUME]}, \texttt{[TRANSFORM]}, \texttt{[BOOST]}, \texttt{[DEBT]} \\
\textbf{Materials:} The complete consumption of a significant item (weapon, book, relic). \\
\textbf{Effect:} Permanently absorb the item's essential nature. Gain one minor property or \textbf{+1 die} to a closely related skill. The item is utterly destroyed. \\
\textbf{Push It:} Take a major property, but gain a permanent \textit{Hunger Mark} (a visible sign of consumption). \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Identity Meal (Standard, 7 XP)} \hfill \\
\emph{Scene; Near; Yes.} \\
\textbf{Tags:} \texttt{[CONSUME]}, \texttt{[TRANSFORM]}, \texttt{[BOOST]}, \texttt{[RISK]} \\
\textbf{Materials:} A personal token or a clear memory of the target. \\
\textbf{Effect:} Temporarily adopt one aspect of a specific person's nature (skill, talent, or defining trait). Gain \textbf{+2 dice} to actions matching that aspect. \\
\textbf{Push It:} The adoption deepens and clings; mark \textbf{1 SB (Hearts)}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).} \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}.

\paragraph{Rite of the Metamorphic Gluttony (High, 13 XP)} \hfill \\
\emph{Scene; Zone; Yes.} \\
\textbf{Tags:} \texttt{[CONSUME]}, \texttt{[AREA]}, \texttt{[TRANSFORM]}, \texttt{[PRESSURE]}, \texttt{[FATE]} \\
\textbf{Materials:} A site scarred by great consumption. \\
\textbf{Effect:} Become a locus of transformative hunger. All present may consume aspects (skills, memories, properties), but each must pass \textit{Spirit+Resolve} (DV 4) or suffer uncontrolled transformation effects. \\
\textbf{Push It:} Some changes become permanent; mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} 1 action; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Hunger's Ascension (High, 14 XP)} \hfill \\
\emph{Extended; Self; Yes.} \\
\textbf{Tags:} \texttt{[TRANSFORM]}, \texttt{[BOOST]}, \texttt{[SACRIFICE]}, \texttt{[FATE]}, \texttt{[DEBT]} \\
\textbf{Materials:} The deliberate offering-up of your current limits to Vorthak. \\
\textbf{Effect:} Consume your own constraints and transform. Choose three: +1 Attribute; +2 dice to one skill; immunity to one Condition type; or one major supernatural capability. \\
\textbf{Push It:} Take all chosen benefits, but lose one core aspect of your original identity; mark \textbf{Harm 2}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons).} \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Obligation:} 8 segments.

% --- Hunger Ascetic (Metamorphic Monk) tcolorbox ---
\begin{tcolorbox}[colback=black!5,colframe=red!70!black,title=\textbf{Hunger Ascetic (Metamorphic Monk)},breakable]
\emph{To consume is to become. To devour is to transform. I am the stomach of the world.}

\vspace{2mm}
\noindent\textbf{Prerequisites:} Body 3+, Spirit 2+, Rite of the Hungry Touch or Cantor of Vorthak.

\vspace{2mm}
\noindent\textbf{Ascetic Techniques (purchased as Talents):}

\begin{tblr}{
  colspec = {X[0.8,l] X[1.8,l] X[1.0,l] X[1.2,l]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Technique & Effect & Cost & Req. \\
\textbf{Gut of Iron} (Basic, 4 XP) & You may consume anything organic (rotten food, raw meat, bitter herbs) without ill effect. +1 die to resist poison or disease from ingestion. & Passive & Body 3 \\
\textbf{Essence Glutton} (Standard, 8 XP) & When you use Rite of the Essence Feast, you may gain one additional minor property (still destroying the item). & +1 Obligation & Tier II \\
\textbf{Hunger's Patience} (Basic, 4 XP) & You may go without food or water for twice as long as normal. When you finally consume, you heal 1 Fatigue. & Passive & Body 3, Spirit 2 \\
\textbf{Metabolic Shift} (Standard, 8 XP) & Once per scene, as a reaction, you may convert 1 Harm you would take into 1 Fatigue (you metabolize the injury). & 1 Boon & Body 4 \\
\textbf{Identity Digestion} (Standard, 9 XP) & When you use Rite of the Identity Meal, the adopted aspect lasts for the entire session instead of one scene. & +1 Obligation & Tier III \\
\textbf{All-Devouring Maw} (Standard, 8 XP) & Once per session, you may consume a single mundane object (size no larger than a shield) as an action, gaining 1 Boon. The object is utterly destroyed. & Free & Body 4, Spirit 3 \\
\end{tblr}

\vspace{2mm}
\noindent\textbf{Master Technique (Epic Talent, 12 XP):}
\textit{Apotheosis of Hunger.} Once per session, you may enter a state of perfect consumption for one scene. You ignore all penalties from Fatigue, and any time you would take Harm from a physical attack, you may instead consume the attack's energy, converting the Harm into 1 Fatigue and gaining +1 die on your next action. When the scene ends, you suffer Harm 2 (Stress) as your body rejects the excess. You cannot use this talent again until you have consumed a significant meal (at least a full day's rations) in a ritual manner.
\end{tcolorbox}

\subsubsection*{Cantors \& Cults of Vorthak}

\textbf{The Cantor's Song.} A Cantor of Vorthak sings in \textit{hungry harmonies} --- melodies that start with a single, clear note and then devour themselves, ending in a hungry silence. Their voice is used to urge a crowd to feast, to sing a blade sharper by absorbing the steel's flaws, or to make a meal last for hours by consuming the very passage of time. Vorthak's Cantors are often found in feasting halls, in butcheries, and around campfires where stories are consumed and retold.

\textbf{The Cult of the Ravenous Circle.} Vorthak's cults meet in places that embody consumption: abattoirs, forges, bakeries, and the kitchens of great manors. The Ravenous Circle is a fellowship of chefs, alchemists, tanners, and smelters --- those who understand that to create something new, something old must be utterly consumed. A ritual of the Circle involves each member bringing a raw material (a cut of meat, a lump of ore, a written secret) and placing it in a central hearth. The hearth consumes it, and the circle speaks aloud what they hope to become by that consumption. Their creed is carved above the hearth: \textit{"What is not consumed is wasted. What is wasted is an insult to hunger."} Outsiders are welcome to bring an offering and a wish, but they must eat a portion of the ash.

\subsubsection*{Witchcraft of Vorthak (Lay / Consumption)}

A witch who walks with Vorthak is a \textit{hunger-smith} or \textit{ash-weaver} --- one who crafts through the act of consumption: a knife that sharpens itself by eating the metal of its victims, a forge that devours impurities and returns only the strongest steel, or a culinary knife that seasons meat by absorbing the memory of the animal's life. Their magic works through the materials of consumption: a quill that writes in ink made from burnt secrets, a needle that sews a garment by consuming thread and producing new cloth, or a smooth stone that absorbs properties by touch. Their tools are never hammers; they use stones, ashes, and the patient heat of a slow digestion.

\textbf{The Witch's Tool: The Maw-Stone.} A smooth, porous stone, warm to the touch, the size of a fist. It is said to have been regurgitated from the belly of a great beast. Once per session, the witch may press the stone against a material (stone, wood, metal, flesh) and speak a word of consumption. The stone ``eats'' a small portion of the material (a handful of stone dust, a sliver of wood, a thin curl of metal) and absorbs its essential property. For the next scene, the stone can be used to grant +1 die to any roll that benefits from that property (e.g., ``sharpness'' for a cutting tool, ``hardness'' for a shield). The stone can also be used to ``consume'' a single, small physical detail (a button, a lock, a hinge) as if it had never been there, but doing so empties the stone of its stored property.

\textbf{A Hedge Gift: Ash of Old Forms (Basic, 4 XP).} Once per scene, the witch may take a handful of ash from a fire that consumed something significant (a letter, a garment, a meal) and blow it into the air. For the next minute, the ash reveals the shape of what was consumed --- a phantom image of the letter, the garment, or the meal. The image is ghostly and cannot be touched, but it can be read, examined, or used to identify an object that no longer exists.

\subsubsection*{Vorthak's Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1.5cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Hunger's Sight:} +1 die to \textit{Notice} consumable properties in objects, beings, or ideas. & \textbf{Metabolic Craving:} Must consume something \emph{significant} daily or suffer \textbf{Fatigue 1}. \\
2 & \textbf{Essence Thief:} Once/scene, temporarily gain a minor property by consuming it. & \textbf{Identity Bleed:} Absorbed aspects surface unbidden; --1 die to maintain your original persona. \\
3 & \textbf{Metamorphic Resilience:} Once/session, consume a \textit{Condition} or \textit{Harm} to negate it. & \textbf{Hunger Aura:} Those Near feel urges to consume or hoard; they suffer --1 die to self-control rolls. \\
4 & \textbf{Transformation Feast:} Once/scene, consume and permanently absorb one \emph{minor} supernatural ability. & \textbf{Essence Instability:} Conflicting properties clash; --1 die when relying on multiple absorbed aspects. \\
5 & \textbf{Identity Cannibal:} Once/session, consume a \emph{significant} aspect of another (major skill or core trait). & \textbf{Self-Consumption:} Risk eating your own memories or abilities; mark \textbf{1 SB (Spades)} when stressed. \\
6+ & \textbf{Hunger Incarnate:} Once/session, become living consumption; absorb nearby properties, abilities, or aspects for one scene. You may consume one physical object as a free action. & \textbf{Void Hunger:} Mark +3 Obligation; risk dissolving into pure appetite that must feed every hour to keep form. Your eyes reflect no light. \\
\end{tblr}

\subsection*{Playstyle Notes}
Vorthak favors those who hunt for change through taking in what the world offers: relish clever uses of absorbed traits, judge the \emph{price} of every bite, and beware the slow surrender of the self. Expect scenes of metamorphosis, bargaining for what can be eaten, and hard choices where power tastes like you. Ideal for characters who embrace transformation at the cost of identity --- alchemists, cannibal cultists, and those who believe that to improve, one must first consume.

\noindent\textbf{Vorthak emphasizes}
\begin{itemize}
  \item \textbf{Transformation through Consumption:} becoming what you devour
  \item \textbf{Metamorphic Power:} fundamental, identity-shaping change
  \item \textbf{Essence Absorption:} discerning and using a thing's core nature
  \item \textbf{Identity Metabolism:} integrating (or failing to integrate) stolen aspects
  \item \textbf{Hunger Propagation:} letting appetite reshape people and places
\end{itemize}