% −−− Patron: The Witness (Truth & Revelation) −−−
\subsection{The Witness (Truth & Revelation)}

\textit{Lore.} The Witness remembers what others bury. Every shadow cast and oath broken is a line in her unending ledger. She is the keeper of inconvenient truths, the patron of those who seek to expose lies or recover forgotten knowledge. Her followers learn that knowledge comes with a price−−−the weight of remembering what others would forget.

The Witness has no will, only a ledger; no mercy, only accuracy. She does not demand confessions, but she ensures that every lie leaves a stain, every secret a trace. Those who serve her become archivists of the uncomfortable, witnesses to the truths that others would kill to keep hidden. To walk her path is to accept that some truths are weapons, and that the truth−teller is never loved.

Among the peoples of Fate's Edge, the Witness is known by many names. In Ecktoria, she is the \textit{Ink‑Eye}, the Censor's clerk who copies every redacted line onto a private scroll. In Vhasia, she is the \textit{Fractured Mirror}, prayed to by spies who seek to reveal a rival's true loyalty. In the Mistlands, she is the \textit{Fog‑Scribe}, whose name is whispered when a bell‑line fails and the dead must be identified by their secrets. And in the shadows of every great city, there are those who follow the \textit{Apocalyptic Witness} – a doomsaying cult that believes the world will end when the last hidden truth is finally spoken aloud, and who work to hasten that revelation by exposing every lie, no matter the cost.

\begin{quote}
I will show you what you would rather forget. But first, you must forget what you think you know.
\end{quote}

\noindent\textbf{Domain Focus}
\begin{itemize}
  ℹtem \textbf{Truth Revelation:} exposing lies, uncovering secrets, forcing confessions
  ℹtem \textbf{Memory Keeping:} preserving knowledge, recalling forgotten facts
  ℹtem \textbf{Inconvenient Knowledge:} learning what others wish to hide
  ℹtem \textbf{Prophetic Warning:} seeing the consequences of hidden truths
  ℹtem \textbf{Burden of Witness:} the psychological weight of knowing too much
\end{itemize}

⫕section*{Rites of the Witness}

\paragraph{Rite of the Lingering Trace (Low, 4 XP)} ℎfill \\
\emph{Instant; Near; Yes (Investigation/Notice only).} \\
\textbf{Tags:} \texttt{[INVESTIGATION]}, \texttt{[TRUTH]} \\
\textbf{Materials:} A trace of the thing to be remembered (hair, dust, a spoken name). \\
\textbf{Effect:} You follow a hidden thread. Gain \textbf{+1 die} to investigate or notice something directly related to the trace within the current scene. Additionally, you may ask the GM one yes/no question about whether the trace is connected to a specific person, event, or object. \\
\textbf{Push It:} Gain \textbf{+2 dice} instead, but mark 1 segment on a \textbf{Memory Strain Timer [4]}. If the timer fills, you gain Fatigue 1 and suffer –1 die on Investigation/Notice rolls until the end of the next scene (the Witness taxes your mind). \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Uncomfortable Question (Low, 5 XP)} ℎfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[COMMAND]}, \texttt{[FEAR]}, \texttt{[TRUTH]} \\
\textbf{Materials:} A circle drawn with chalk or string while focusing on a specific person. \\
\textbf{Effect:} You force a moment of raw honesty. Ask one question of a target within the zone. They may either answer truthfully (the GM must give a true answer, though it may be cryptic) or mark \textbf{1 Fatigue} and gain a \textit{Guilty Tell} Condition ( –1 die to all social rolls until they confess the truth to someone). \\
\textbf{Push It:} The target must also reveal a surface emotion or a forgotten detail (GM’s choice). Mark \textbf{1 SB (Hearts)} as the Witness records the strain. \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Echoing Truth (Standard, 8 XP)} ℎfill \\
\emph{Instant; Near; No.} \\
\textbf{Tags:} \texttt{[READ]}, \texttt{[TRUTH]}, \texttt{[FEAR]} \\
\textbf{Materials:} A reflective surface (mirror, still water, polished metal) used to focus on the target. \\
\textbf{Effect:} The target must make a Resolve test (DV 3) or suffer –1 die to rolls involving memory, deception, or resisting interrogation for the scene. If they fail, you may ask one specific, factual question about something they know, and they must answer truthfully or suffer \textbf{1 SB (Hearts)} as the memory is forcibly drawn forth. \\
\textbf{Push It:} If the target fails their Resolve test, you may ask a second question, but the mental intrusion causes them \textbf{Harm 1 (Stress)}. \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Immutable Record (Standard, 7 XP)} ℎfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[VOW]}, \texttt{[BOND]}, \texttt{[TRUTH]} \\
\textbf{Materials:} A document signed by all parties within the zone, or a spoken pact witnessed by the caster. \\
\textbf{Effect:} Bind the agreement. Any party who knowingly breaches it suffers \textbf{1 SB (Hearts)} immediately and gains a persistent \textbf{Oathbreaker’s Mark} Condition (–1 die on social rolls involving honour, trust, or oaths until amends are made or a significant act redeems them). \\
\textbf{Push It:} The bond becomes magically enforced for one specific, crucial clause: violation automatically inflicts \textbf{Harm 1 (Stress)} on the breaker in addition to the SB and Mark. \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Shared Burden (Standard, 9 XP)} ℎfill \\
\emph{Instant; Touch; No.} \\
\textbf{Tags:} \texttt{[HEAL]}, \texttt{[DEBT]}, \texttt{[READ]} \\
\textbf{Materials:} A written secret (yours or another’s) and a candle. \\
\textbf{Effect:} You transfer the weight of a truth. Choose one: you may remove a \textit{Guilty Tell}, \textit{Haunted}, or similar psychological Condition from yourself and give it to a willing or unwilling target (they resist with Resolve DV 4). Alternatively, you may share a secret with an ally; both of you clear 1 Fatigue as the burden is halved. \\
\textbf{Push It:} Transfer a Condition without the target’s consent (they automatically resist, but on failure they gain it). Mark \textbf{1 SB (Spades)} as the secret stains them. \\
\textbf{Invoke:} 1 action; mark \textbf{+1 Obligation}. \\
\emph{Requires: Familiar + Codex (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Unveiled Heart (High, 12 XP)} ℎfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[MANIPULATION]}, \texttt{[FEAR]}, \texttt{[TRUTH]}, \texttt{[AREA]} \\
\textbf{Materials:} A private setting where the target feels safe or is speaking freely. \\
\textbf{Effect:} The target suffers –2 dice to all attempts to conceal true emotions, intentions, or lies for the scene. Any successful social roll (Sway, Command, Deceit) made by the target generates \textbf{1 SB (Hearts)} as the effort to maintain falsehoods under the Witness’s gaze creates internal discord. \\
\textbf{Push It:} You may designate one specific, complex question about the target’s motivations, fears, or hidden loyalties. If you successfully use Sway or Insight against them this scene, you automatically learn the answer to that question. The intense scrutiny marks \textbf{1 SB (Spades)} for you as the Witness’s attention lingers. \\
\textbf{Invoke:} 1 action; mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 6 segments.

\paragraph{Rite of the Final Reckoning (High, 13 XP)} ℎfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[COMMAND]}, \texttt{[TRUTH]}, \texttt{[FEAR]} \\
\textbf{Materials:} A formally called gathering (court, council, family meeting) within the consecrated zone. \\
\textbf{Effect:} All present must speak their greatest debt, wrongdoing, or hidden truth related to the gathering’s purpose. Those who lie or withhold suffer \textbf{Harm 2 (Stress/Reputation)}. Truth‑tellers gain \textbf{+2 dice} to social actions for the remainder of the scene within the zone. \\
\textbf{Push It:} The truth becomes inescapable – even indirect lies or evasions related to the core topic suffer the Harm 2 penalty. The absolute nature of the revelation creates \textbf{2 SB (Diamonds)} as the disruption to fates and secrets resonates. \\
\textbf{Invoke:} Extended rite; mark \textbf{+2 Obligation}. \\
\emph{Requires: Familiar + Codex + Tier III (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

⫕section*{Cantors & Cults of the Witness}

\textbf{The Cantor’s Song.} A Cantor of the Witness sings in \textit{truth‑tolling} – clear, piercing melodies that cut through evasion, harmonies that make the hidden surface. Their voice is used to compel a confession, to sound an alarm that a secret is about to be spoken, or to sing a historical record into being, each note a word in a ledger. The Witness’s Cantors are often found in courtrooms, in archive vaults, and at the bedsides of the dying, singing the last truth.

\textbf{The Congregation of the Unmasking (Apocalyptic Cult).} The most feared of the Witness’s cults, the Congregation of the Unmasking believes that the world will end when the last hidden truth is finally revealed – and that this ending is not a catastrophe but a liberation. They seek to expose every lie, every buried crime, every forgotten promise, no matter the cost. A ritual of the Congregation involves each member speaking aloud a truth they have kept secret, then burning a representation of the lie that concealed it. Their creed is painted on a scroll that is never fully unrolled: \textit{“When all that is hidden is seen, the world will either break or be free. We will be there to witness which.”} Outsiders are not welcome; they are potential sources of truth – or obstacles to be unmasked. The Congregation has been known to destroy families, topple governments, and start wars in their pursuit of total revelation. They are loved by none and feared by all, but even their enemies cannot deny that they are right: some secrets truly are worth dying to expose.

⫕section*{Witchcraft of the Witness (Craft / Record)}

A witch who walks with the Witness is a \textit{ledger‑scribe} or \textit{seal‑breaker} – one who crafts the tools of permanent record: a book that cannot be altered once written, a quill that writes in ink that fades only when the truth it records is fulfilled, or a press that stamps each page with a sigil of authenticity. Their magic works through the materials of documentation: a chisel that carves a truth into stone so that it cannot be erased, a needle that sews a confession into a garment so that the wearer cannot deny it, or a forge that tempers a door that can only be opened by someone who has never told a lie. In Aeler, a True‑Mason might carve a “witness stone” at a courthouse entrance, so that all testimony given within is remembered by the stone for a century. In Thepyrgos, a bookbinder might stitch a volume that screams if its pages are cut.

\textbf{The Witch’s Tool: The Quill of Unwritten Truths.} A quill carved from a raven’s feather, its nib dipped in ink made from lampblack and a drop of the witch’s blood. Once per session, the witch may write a single sentence on a piece of parchment, then seal it with a drop of wax. For the next day, anyone other than the witch who reads the sentence must make a Resolve test (DV 3) or forget it immediately after reading. The witch may also use the quill to “seal” a spoken secret, causing the speaker’s voice to catch if they try to repeat it for the rest of the scene.

\textbf{A Hedge Gift: Record of a Breath (2 XP).} Once per scene, the witch may hold a blank piece of parchment up to a person’s lips as they speak. The parchment absorbs the spoken words, recording them exactly as they were said – including tone, hesitation, and any unconscious meaning. The parchment can be read by anyone (the words appear in the speaker’s own handwriting), but it crumbles to dust after being read aloud. The witch may also use this ability to record the final words of the dying, preserving them for one day.

⫕section*{The Witness’s Corruption Table}

\begin{tblr}{
  colspec = {Q[c,1.5cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Truth’s Sight:} +1 die to Insight when detecting deception or hidden motives. & \textbf{Burden of Knowledge:} –1 die to social rolls involving lies or deception; others become uncomfortable with your piercing gaze. \\
2 & \textbf{Memory’s Keeper:} Once/scene, recall one specific detail from a previous scene with perfect clarity. & \textbf{Compulsive Honesty:} Must correct obvious falsehoods witnessed, even when tactically disadvantageous. \\
3 & \textbf{Revelation’s Power:} +2 dice to rolls involving exposing secrets, uncovering lies, or forcing confessions. & \textbf{Truth‑Blind:} Suffer 1 Fatigue when exposed to comforting lies or wilful ignorance. \\
4 & \textbf{Witness’s Authority:} Once/session, force one target to make a Resolve test (DV 4) or reveal a significant hidden truth. & \textbf{Isolation:} –1 die to rolls requiring trust or close relationships; others fear your ability to uncover their secrets. \\
5 & \textbf{Omniscient Gaze:} Once/session, see through all deceptions and lies for one exchange, gaining +3 dice to related actions. & \textbf{Paranoia:} –1 die to rolls involving personal peace or rest; the weight of all truths witnessed creates constant mental strain. \\
6+ & \textbf{Absolute Witness:} Once/session, become the living embodiment of truth. For one scene, all deceptions within Near range automatically fail. & \textbf{Truth’s Prison:} Mark +3 Obligation when using this power; become unable to tolerate any form of deception, making normal social interaction nearly impossible. \\
    &    & \textbf{The Unmasking Beckons:} The apocalyptic urge grows – you may be called to expose a truth that will destroy a community, a nation, or yourself. \\
\end{tblr}

\subsection*{Playstyle Notes}
The Witness rewards curiosity, patience, and a willingness to bear uncomfortable knowledge. Her followers are not merely investigators but \textit{archivists of the real} – they record what others would erase. The Witness does not demand that you expose every secret; she simply ensures that you cannot forget the ones you learn. Intrusions often take the form of dreams showing a hidden crime, a sudden memory of a lie you told years ago, or the appearance of a scrap of parchment containing a truth you wish you had never read. High‑Tier followers may find that secrets begin to \textit{seek them out} – people confess unprompted, documents fall open to damning pages, and the world becomes a book that cannot be closed.

\noindent\textbf{Emphasises}
\begin{itemize}
  ℹtem \textbf{Truth as Burden:} every secret known is a weight carried
  ℹtem \textbf{Investigation as Ritual:} uncovering lies is a sacred act
  ℹtem \textbf{Memory as Weapon:} forgetting can be as powerful as remembering
  ℹtem \textbf{The Unmasking:} exposing a truth can be an act of mercy or destruction
  ℹtem \textbf{Witness as Isolation:} those who know too much are seldom loved
\end{itemize}

% −−− Monastic Tradition: Unmasked Fist (The Witness) −−−
\begin{tcolorbox}[colback=black!5,colframe=yellow!70!black,title=\textbf{Unmasked Fist (The Witness)}]
  \emph{I will show you what you would rather forget.}
  
  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Spirit 3+, Wits 2+, Rite of the Lingering Trace or Cantor of Witness.
  
  \vspace{2mm}
  \noindent\textbf{Debt–Resistant Frame.} The Witness records, but does not collect. Monks of the Witness gain \textbf{+1 die} to resist any magical debt or obligation that would force them to conceal the truth or perjure themselves. The creditor’s ledger will show a blank where their name should be.
  
  \vspace{2mm}
  \noindent\textbf{Techniques:}
  
  \begin{tblr}{
    colspec = {X[0.6,l] X[1.8,l] X[1.2,l] X[1.4,l]},
    rowhead = 1, row{odd}={bg=gray!10}, row{1}={bg=gray!30}, hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Confession Blow} & On an unarmed hit, ask the GM one yes/no question about the target’s hidden deed or recent lie. The GM must answer truthfully. & 1 Boon & Spirit 3 \\
  \textbf{Lingering Trace} & Follow a hidden thread. Gain +1 die to investigate a specific trace within the current scene. & 1 Fatigue & Wits 2 \\
  \textbf{Uncomfortable Question} & Force a target to answer one question truthfully (GM answer) or mark 1 Fatigue and gain a Guilty Tell (–1 die social rolls). & 1 Boon & Presence 2 \\
  \textbf{Immutable Record} & Bind an agreement. Any party who knowingly breaches it suffers 1 SB (Hearts) and gains an Oathbreaker’s Mark (–1 die social rolls involving honour until amends). & 1 Fatigue & Tier II \\
  \textbf{Shared Burden} & Transfer a psychological Condition (Guilty, Haunted) from yourself to a target (Resist DV 4) or share a secret with an ally to clear 1 Fatigue each. & 1 Boon & Spirit 3 \\
  \end{tblr}
  
  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Absolute Witness.} Once per session, become living truth for one scene: all deceptions within Near range automatically fail. You may demand a single truthful answer to one question from any creature, and they cannot refuse (no resistance roll). After the scene, you gain a permanent \textbf{Debt Scar} (your eyes reflect light like a brass mirror) and may ignore one debt or obligation that would require you to lie – the creditor sees only the truth of your refusal.
\end{tcolorbox}