% ============================================================================
% PATRON ENTRY: XHAK'THUL, THE THUNDER-SPEAKER (STORM, SHOCK & AWE)
% ============================================================================

\subsection{Xhak'Thul, the Thunder-Speaker (Storm, Shock \& Awe)}

\subsubsection*{Lore}
Before words, there was thunder. Before scripture, lightning carved its commandments into the bones of the world. Xhak'Thul is the Voice That Splits Mountains, the Drum That Shakes Foundations, a power born from the first storm that ever raged across the primal earth. His faithful are whirlwinds of action --- they do not study; they dance, they shout, they move with the sky's fury. His gifts are not gentle miracles but raw terror, awe, and vitality ripped from nature's heart.

Among the peoples of Fate's Edge, Xhak'Thul is the patron of the pre-literate, the neolithic, those who have not yet built walls against the wild. In the deep Vhasian highlands, he is the \textit{Storm-Father}, invoked by herders who watch lightning split the sky. In Ykrul, he is the \textit{Thunder-Hoof}, the sound of a thousand horses stampeding. In Linn, he is the \textit{Raging Tide}, the storm that comes without warning. His followers are not scholars; they are dancers, drummers, shamans, and warriors who feel the storm in their bones before they hear it in the sky. He favors those who act without hesitation, who embrace chaos, and who understand that the loudest voice is often the truest.

\begin{quote}
``Feel the drum in your bones before you hear it in the sky. When the storm speaks, you will already be dancing to its rhythm.''
\end{quote}

\subsubsection*{Domain Focus}
\begin{itemize}
  \item \textbf{Primal Power:} storm, thunder, lightning, earthquake, wildfire
  \item \textbf{Shock and Awe:} overwhelming sensory impact, terror as a weapon
  \item \textbf{Fear-Eater:} absorbing panic to fuel strength
  \item \textbf{Immediate Magic:} no rituals, no books --- only emotion, movement, and the moment
\end{itemize}

\subsubsection*{Thiasos and Codex (Runekeeper Options)}

A Runekeeper sworn to Xhak'Thul does not bind a quiet familiar. Their \textbf{Thiasos} is a small beast or mythological creature that embodies the storm's wild heart: an electric blue salamander that crackles when touched, a tiny hawk whose feathers hum with static, a mouse that drums its tail in thunderstorms, or a small snake that vibrates like a plucked string before lightning strikes. This creature requires no food, but it must feel the open sky and hear thunder each week --- a storm, a war drum, or even a slammed door. If neglected for a week, it becomes \textit{Compromised} (it sparks uncontrollably, imposing --1 die on Stealth rolls for its bearer and occasionally setting small fires).

The \textbf{Codex} of Xhak'Thul is never a book of quiet study. It may be a drumhead painted with storm-runes, a set of carved bones that rattle like thunder when shaken, a stone tablet scored by a lightning strike, or a bundle of twigs tied with sinew that snaps like a whip crack when a Rite is invoked. Because the storm resists stillness, the Codex requires upkeep: the Runekeeper must spend an hour each week beating a rhythm into it (or exposing it to loud percussion), or the Rites stored there become sluggish (treat as Compromised until re-awakened).

\subsubsection*{Patron's Gift (Imbuement)}

Once per scene as an action, a Runekeeper of Xhak'Thul may touch a held item (weapon, shield, or drum) and whisper a thunder-name. For the rest of the scene, the item grants \textbf{+1 die} to \textit{Intimidation} or \textit{Command} rolls, and the first time each scene the bearer would be surprised or lose initiative, they may act as if they were not surprised (the storm warns them). After the scene, the user marks \textbf{+1 Obligation} to Xhak'Thul.

\subsubsection*{Rites of Xhak'Thul}

\paragraph{Rite of the Thunderous Presence} (Basic, 4 XP) \hfill \\
\emph{Scene; Near; No.} \\
\textbf{Tags:} \texttt{[FEAR]}, \texttt{[COMMAND]}, \texttt{[AREA]}, \texttt{[PRESSURE]} \\
\textbf{Materials:} A drum made from storm-felled wood or any suitable improvised percussion. \\
\textbf{Effect:} Your presence becomes overwhelming. All NPCs within Near range must test Resolve (DV 2) or suffer --1 die to social actions due to awe or terror for the scene. Allies gain +1 die to Intimidation or Command actions. \\
\textbf{Push It:} Extend the effect to Far range, but you mark 1 Fatigue and mark \textbf{1 SB (Hearts)} as your presence becomes difficult to control. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Storm's Step} (Basic, 5 XP) \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[SPEED]}, \texttt{[FEAR]}, \texttt{[BOOST]} \\
\textbf{Materials:} Earth from a lightning strike site, worn boot leather. \\
\textbf{Effect:} You move with the storm's speed. Gain \textbf{+2 dice} to Athletics or Stealth rolls. The first attack you make this scene gains \textbf{+1 Effect}. \\
\textbf{Push It:} Gain \textbf{+3 dice} and move as a blur, becoming difficult to track or clearly remember; mark \textbf{1 SB (Diamonds)}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Shattered Sky} (Advanced, 8 XP) \hfill \\
\emph{Scene; Zone; No.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[FORCE]}, \texttt{[FEAR]}, \texttt{[WEATHER]} \\
\textbf{Materials:} A shard of meteoric iron or metal touched by a severe storm. \\
\textbf{Effect:} You call a localized, supernatural storm in a Zone. Choose one effect:
\begin{itemize}
  \item \textbf{Lightning Strike:} Single target, Harm 2.
  \item \textbf{Wind Blast:} Knockback, dispersing arrows or scattering foes.
  \item \textbf{Hail Barrage:} Harm 1 (AREA).
\end{itemize}
All enemies in the Zone suffer --1 die to actions from chaos and distraction. \\
\textbf{Push It:} Double the chosen effect's scale or potency (e.g., Harm 3, large knockback), but the effect also applies to allies in the Zone; mark \textbf{2 SB (Clubs)}. Cost: Mark +1 Obligation (total +3) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +2 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the Fear-Eater} (Advanced, 9 XP) \hfill \\
\emph{Scene; Self; No.} \\
\textbf{Tags:} \texttt{[FEAR]}, \texttt{[HEAL]}, \texttt{[COMMAND]}, \texttt{[BOOST]} \\
\textbf{Materials:} Soil or dust from a place of great terror (battlefield, prison, ruins). \\
\textbf{Effect:} You feed on the fear around you. For each Frightened target in the scene, gain \textbf{+1 die} on your next action (max +3). If you successfully Intimidate a target during this scene, you heal 1 Fatigue. \\
\textbf{Push It:} Absorb all palpable fear in the scene, gaining a flat \textbf{+3 dice} bonus to your next action, but mark 1 permanent Scar related to fear. Mark \textbf{1 SB (Hearts)}. Cost: Mark +1 Obligation (total +2) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +1 Obligation. \\
\textit{Requires:} Familiar + Codex (\textit{Invoke:} 1 Boon). \\
\textbf{Invoke:} 1 action. \\
\textbf{Duration:} Scene.

\paragraph{Rite of the World's Drum} (Master, 14 XP) \hfill \\
\emph{Extended; Zone; No.} \\
\textbf{Tags:} \texttt{[AREA]}, \texttt{[FEAR]}, \texttt{[COMMAND]}, \texttt{[TRANSFORM]}, \texttt{[WEATHER]} \\
\textbf{Materials:} A drum fashioned from lightning-struck wood, with a hide from a beast killed by natural force. \\
\textbf{Effect:} You become an incarnation of the storm for one scene:
\begin{itemize}
  \item Gain \textbf{+2 dice} to rolls based on strength, speed, and intimidation.
  \item Your anger echoes as thunder; attacks against you may be answered by lightning (GM decides, usually 1 SB and Harm 1 to the attacker).
  \item Enemies must test Resolve at --2 dice to resist Fear effects you create.
\end{itemize}
Start a 6-segment \textit{Natural Disasters} timer; when it fills, the storm's wrath manifests as a random environmental hazard (landslide, lightning strike, flash flood). \\
\textbf{Push It:} The transformation lingers after the scene, and your emotional state causes minor environmental effects (windows rattle, small fires flicker). Mark \textbf{2 SB (Diamonds)} and advance the timer by 2 segments. Cost: Mark +1 Obligation (total +4) or Mark +1 Fatigue. \\
\textbf{Cost:} Mark +3 Obligation. \\
\textit{Requires:} Familiar + Codex + Tier III (\textit{Invoke:} 2 Boons). \\
\textbf{Invoke:} Extended action. \\
\textbf{Duration:} Scene.

\subsubsection*{Cantors \& Cults of Xhak'Thul}

\textbf{The Cantor's Song.} A Cantor of Xhak'Thul does not sing melodies --- they drum. Their voice is a percussive roar, a heartbeat rhythm that shakes the ground. Their ``song'' is the pattern of a storm: the slow build of pressure, the sudden crack of lightning, the fading rumble of retreating thunder. Xhak'Thul's Cantors are often found in places without writing, where memory is carried in dance and the beat of a skin drum.

\textbf{The Cult of the First Drum.} Xhak'Thul's cults gather in clearings, on hilltops, or in caves that echo with thunder. The First Drum is a fellowship of dancers, shamans, and warriors who believe that civilization is a cage and that the storm is the only honest law. A ritual of the Cult involves each member drumming on a skin drum (never bought, always made from an animal they killed themselves) until the first person collapses from exhaustion. The collapsed dancer is declared ``storm-taken'' and must lead the next ritual. Their creed is not written --- it is beaten into the earth: \textit{``The sky does not ask permission. Neither do we.''} Outsiders are welcome to dance, but they must bring a drum they made with their own hands.

\subsubsection*{Witchcraft of Xhak'Thul (Lay / Storm)}

A witch who walks with Xhak'Thul is a \textit{storm-caller} or \textit{drum-maker} --- a shaman who crafts the tools of primal power: a drum that can be heard for a mile, a spear of lightning-struck wood that never misses its target, or a charm of meteoric iron that calls a storm when struck. Their magic is immediate, visceral, and never subtle. In Ykrul, a bone-carver might shape a whistle that mimics the sound of a thunderclap, terrifying horses. In the deep Vhasian highlands, a shaman might paint their drum with the blood of a goat struck by lightning, binding the storm's spirit to the skin.

\textbf{The Witch's Tool: The Storm-Drum.} A hand-drum made from lightning-struck wood and the hide of an animal killed by a natural disaster (flood, fire, landslide). The drum's surface is painted with a spiral that represents the eye of the storm. Once per session, the witch may strike the drum with a single, thunderous beat. All creatures within Near range must test Resolve (DV 3) or be Shaken (--1 die to all actions) for one exchange. The drum can also be used to ``call the storm'' --- beating a steady rhythm for a full minute (out of combat) grants +1 die to all weather-related rolls for the next hour.

\textbf{A Hedge Gift: Drum of the First Beat (Basic, 4 XP).} Once per scene, the witch may tap a rhythm on any surface (a table, a shield, a tree trunk). For the next minute, all allies within Near range who hear the rhythm gain \textbf{+1 die} to Athletics and combat rolls, as their heartbeat synchronizes with the storm. The rhythm cannot be sustained if the witch is silenced or takes a wound.

\subsubsection*{Corruption of Xhak'Thul}

\begin{tblr}{
  colspec = {Q[c,1cm] X[2.5] X[2.5]},
  rowhead = 1,
  row{odd} = {bg=gray!10},
  row{1} = {bg=gray!30, font=\bfseries},
  hlines,
}
Tier & Benefit & Cost / Quirk \\
1 & \textbf{Static Cling:} Your touch carries a static shock. Metal objects faintly hum in your presence. & You feel restless energy before a storm. Minor lies itch like a rash. \\
2 & \textbf{Storm-Touched:} Your hair stirs without wind. Your voice occasionally cracks like thunder. Small flames flicker near you. & You become irritable in calm, silent places. You crave loud arguments and violent weather. \\
3 & \textbf{Lightning Rod:} You are dangerously magnetic. Unshielded metal flies to your hands. Your eyes crackle with faint light. & You feel a compulsion to disrupt order and create chaos. Peace feels like a cage. \\
4 & \textbf{Walking Storm:} A localized wind stirs around you. You smell of ozone and wet stone. You cannot whisper; your voice is a low rumble. & You lose patience with lengthy plans. You prefer direct, explosive action, regardless of the cost. \\
5 & \textbf{Heart of the Tempest:} You are the eye of a personal storm. Rain sheets indoors; lightning threatens those who anger you. & You see the world only as a force to be broken or commanded. Gentler emotions are foreign memories. \\
6+ & \textbf{Living Thunder:} Once/session, you may become a bolt of lightning, teleporting to any point you can see within Far range and unleashing an automatic Harm 2 blast upon arrival. & \textbf{Storm's Prison:} Mark +3 Obligation. You can no longer remain indoors for more than an hour without suffering 1 Fatigue; the sky demands your presence. \\
\end{tblr}

\subsection*{Playstyle Notes}
Xhak'Thul is the patron of shamans, berserkers, and anyone who acts first and thinks never. His magic is immediate, loud, and unsubtle --- the thunderclap, the lightning strike, the roar that silences a crowd. He rewards boldness, spontaneity, and the willingness to break things. His followers are not subtle, but they are terrifying. Ideal for players who enjoy direct solutions, primal power, and characters who would rather break a door than pick a lock.

\noindent\textbf{Emphasizes}
\begin{itemize}
  \item \textbf{Primal Power:} storm, thunder, lightning, earthquake, wildfire
  \item \textbf{Shock and Awe:} overwhelming sensory impact, terror as a weapon
  \item \textbf{Fear-Eater:} absorbing panic to fuel strength
  \item \textbf{Immediate Magic:} no rituals, no books --- only emotion, movement, and the moment
  \item \textbf{Pre-Literate Wisdom:} the oldest magic, written in bone and drum
\end{itemize}

% --- Monastic Tradition: Storm Fist (Xhak'Thul) ---
\begin{tcolorbox}[colback=black!5,colframe=cyan!50!black,title=\textbf{Storm Fist (Xhak'Thul)},breakable]
  \emph{The storm does not ask. It speaks. You are its echo.}

  \vspace{2mm}
  \noindent\textbf{Prerequisites:} Body 3+, Spirit 2+, Rite of the Thunderous Presence or Cantor of Xhak'Thul.

  \vspace{2mm}
  \noindent\textbf{Debt-Resistant Frame.} Lightning leaves no ledger. Monks of the Storm Fist gain \textbf{+1 die} to resist any effect that would bind, contract, or obligate them (geas, oath, debt-enforcement). If they succeed, the air around them crackles audibly --- pursuers gain +1 die to track them for the next hour.

  \vspace{2mm}
  \noindent\textbf{Techniques (purchased as Talents):}

  \begin{tblr}{
    colspec = {X[0.7,l] X[1.8,l] X[1.1,l] X[1.4,l]},
    rowhead = 1,
    row{odd}={bg=gray!10},
    row{1}={bg=gray!30, font=\bfseries},
    hlines,
  }
  Technique & Effect & Cost & Req. \\
  \textbf{Thunder Palm} (Basic, 5 XP) & Unarmed strike deals +1 Harm and forces a Resolve test (DV 3) or the target loses their next minor action. & 1 Fatigue & Body 3 \\
  \textbf{Lightning Reflexes} (Basic, 4 XP) & +1 to defense DV against ranged attacks. & Passive & Body 3, Athletics 2 \\
  \textbf{Fear Eater} (Basic, 5 XP) & When you successfully Intimidate a target, gain 1 Boon (once per scene). & Passive & Spirit 3, Command 2 \\
  \textbf{Storm Step} (Advanced, 8 XP) & Move as a blur; you are difficult to track or remember. For one exchange, stealth rolls are at +2 dice. & 1 Boon & Spirit 3, Tier II \\
  \textbf{Storm's Pace} (Basic, 4 XP) & While moving (not standing still), gain +1 die to all physical actions. & Passive & Body 3, Performance 1 \\
  \textbf{Thunder's Grasp} (Advanced, 9 XP) & When reduced to 0 Harm, instead unleash a concussive blast: all enemies in Close are knocked back one band and must test Body+Resolve (DV 3) or fall prone. You survive with Fatigue 2. & Once per session & Tier III, Fear Eater \\
  \end{tblr}

  \vspace{2mm}
  \noindent\textbf{Master Technique (Epic Talent, 12 XP):}
  \textit{Eye of the Storm.} Once per session, become the calm center of a supernatural storm for one scene. Allies within Near ignore fear and weather penalties. You gain +2 dice to physical and intimidation rolls, and any enemy that attacks you suffers a lightning counter (Harm 1, ignores armor). After the scene, you collapse with Fatigue 3. Gain a permanent \textbf{Debt Scar} --- a lightning-bolt pattern on your chest --- but you may ignore one forced Obligation or debt call (the creditor sees only static).
\end{tcolorbox}