\subsubsection{Grimmir, the Old Man of the Forest (Instinct \& Wild Memory)}
\begin{quote}\itshape
“They say he taught the wolves to stalk, and the children of men to kindle fire. Listen close, and you may remember what your blood already knows.”
\end{quote}

\paragraph{Lore.}
In mist-thick woods and hollow caves, the stories whisper of \textbf{Grimmir}, the Old Man of the Forest. Neither wholly human nor wholly beast, he is the memory of the primal hunter—matted hair, claw-like nails, the antlers of a stag and the eyes of a wolf. He is not a god, but a reminder: that humankind was once feral, and the wild still claims its due. Those who call on him gain the instincts of predator and prey alike, yet risk losing their civil tongue for the growl and the howl.

\paragraph{Patron’s Gift — Beastial Relic.}
Your sworn implement (club, bone spear, or totem mask) sprouts hair, fang, or horn.
\begin{itemize}
  \item \textbf{Enchanted Weapon:} Counts as enchanted melee \textbf{+1 Melee}.
  \item \textbf{Thematic Skill:} You gain \textbf{+1 to Intimidation or Survival} when drawing on primal instinct.
\end{itemize}

% =========================
% RITES
% =========================

\paragraph{Hunter’s Snarl (Low, 4 XP)} \emph{Scene; Self; Yes.}\\
\textbf{Materials:} A scrap of bone or a beast’s tooth.\\
\textbf{Effect:} Bare teeth, snarl, and draw on animal menace. Gain +1 die to Intimidation or resist fear this scene.\\
\textbf{Push It:} Enemies hesitate—first hostile act against you suffers --1 die.\\
\textbf{Backlash (Fate):} Your instincts flare—--1 die on attempts at polite social interaction this scene.

\paragraph{Wolf-Step (Low, 4 XP)} \emph{Scene; Self; Yes.}\\
\textbf{Materials:} Dirt from a wolf’s track or claw scratch.\\
\textbf{Effect:} Move with lupine cunning. Gain +1 die to Stealth or Athletics when flanking or stalking.\\
\textbf{Push It:} Also ignore one minor environmental penalty (brambles, roots).\\
\textbf{Backlash (Body):} Predatory gait aches your joints—minor Fatigue until you rest.

\paragraph{Horn of the Wild Hunt (Standard, 8 XP)} \emph{Scene; Near; Yes.}\\
\textbf{Materials:} A horn, conch, or hollowed bone.\\
\textbf{Effect:} Sound the primal call. Allies gain +1 die to their next attack or chase roll this scene. Enemies of lesser courage must resist or falter.\\
\textbf{Push It:} Also call a Cap 1 beast follower (a hound, crow, or stag) until the scene ends.\\
\textbf{Backlash (Spirit):} The horn echoes back; GM gains 1 SB to spend on attracting a rival predator.

\paragraph{Mark of Antlers (Standard, 8 XP)} \emph{Scene; Self or Touch; Yes.}\\
\textbf{Materials:} Ash or blood traced across the brow.\\
\textbf{Effect:} Sprout phantom antlers and command primal space. Gain +1 die to Command or resist physical shoves; foes treat you as larger.\\
\textbf{Push It:} For one exchange, enemies in Near range must test or suffer --1 die to strike you.\\
\textbf{Backlash (Life):} The mark lingers—--1 die to Diplomacy checks until it fades.

\paragraph{Wild Communion (Major, 12 XP)} \emph{Scene; Region; Yes.}\\
\textbf{Materials:} Sit upon raw earth, surrounded by a circle of bones.\\
\textbf{Effect:} Commune with primal memory. Ask two questions about the movement of beasts, predators, or prey in the last day. Gain advantage on the next hunt, track, or ambush.\\
\textbf{Push It:} Ask a third question, or extend your communion to human hunters nearby.\\
\textbf{Backlash (Dreams):} You wake that night with wild hunger—GM may introduce an omen or Complication.

\paragraph{Shape of the Ancestor (Epic, 16 XP)} \emph{Scene; Self; Yes.}\\
\textbf{Materials:} Wear a mask of fur, antler, or bone.\\
\textbf{Effect:} Transform into a primal hybrid—fangs, claws, heavy pelt. For the scene:
\begin{itemize}
  \item +1 Body
  \item Natural melee attack (enchanted, +1)
  \item +1 die to Athletics, Melee, or Survival
\end{itemize}
\textbf{Push It:} Extend to a second ally, marking them with blood or ash.\\
\textbf{Backlash (Life/Death):} When it fades, you suffer a minor Exhausted condition, and the Old Man whispers in dreams.
