% patrons/khemesh.tex
% Fate’s Edge — Patron: Khemesh, the Abyssal Maw

\subsubsection{Khemesh, the Abyssal Maw (Depths, Inexorability, Eldritch Terror)}
\textit{Lore.} Khemesh is not merely a lord of the depths but the hunger beneath them, a pressure older than seas. Those who bargain with him are marked by the abyss—seen in the way shadows cling, in the whispers heard when no voice speaks, in the certainty that all things will sink.

\begin{quote}
In the trench without light, the Maw waits. Even silence drowns.
\end{quote}

\paragraph{Whisper of the Trench (Low, 4 XP)} \emph{Instant; Near; No.}\\
\textbf{Effect:} Target hears impossible echoes and suffers \textbf{−1 die} on their next action.\\
\textbf{Push It:} Echoes coil in your own skull—take \textbf{Fatigue 1}, but the target also loses their next minor action.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of Crushing Silence (Low, 5 XP)} \emph{Scene; Zone; No.}\\
\textbf{Materials:} A broken shell filled with ink-dark water.\\
\textbf{Effect:} Establish an oppressive silence; sound carries only as distorted whispers. Enemies in the zone gain \textbf{−1 die} to coordination or morale-driven actions.\\
\textbf{Push It:} A single enemy’s voice is stolen entirely for the scene.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Pressure of the Maw (Standard, 7 XP)} \emph{Instant; Near; No.}\\
\textbf{Materials:} A length of rusted chain submerged in water.\\
\textbf{Effect:} Target is pinned by invisible crushing force: treat as \texttt{[ENTANGLE]} with \textbf{Great Effect} if underwater or confined.\\
\textbf{Push It:} Inflict \textbf{Fatigue 1} on the target in addition to the restraint.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Abyssal Vision (Standard, 9 XP)} \emph{Scene; Self; No.}\\
\textbf{Effect:} You perceive the world as Khemesh does—fractured, alien, crushing. Gain \textbf{+2 dice} to Notice and Arcana, and may ask one “true nature” question about a foe or structure.\\
\textbf{Cost:} When the scene ends, you suffer \textbf{Exposure +1} as your perception warps.\\
\textbf{Push It:} Extend the vision to one ally, but both take \textbf{Fatigue 1}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{The Maw Opens (High, 12 XP)} \emph{Scene; Zone; No.}\\
\textbf{Materials:} A sealed vessel of abyssal water, broken open.\\
\textbf{Effect:} Reality in the zone folds inward like the crushing deep: \\
\begin{itemize}
  \item Enemies act at \textbf{Desperate Position} by default.  
  \item Each beat, the Keeper may force \textbf{1 SB} (Spades/Clubs favored).  
  \item Structures, vessels, or wards fracture as if under immense weight.  
\end{itemize}
\textbf{Push It:} For one beat, declare a single enemy “crushed” (severe harm/effect). You immediately suffer \textbf{Fatigue 2} and \textbf{+1 Obligation}.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} 2 Boons).}\quad \emph{Obligation:} 8 segments.

\paragraph{Rivalry: Raéyn.} Khemesh embodies the unknowable trench that swallows sailors; Raéyn embodies tides, travel, and the sea’s surface. Where Raéyn charts and protects, Khemesh unmoors and devours. In scenes of open sea, Raéyn gains the upper hand; in the abyssal dark, Khemesh dominates.
