% patrons/nidhoggr.tex
% Fate’s Edge — Patron: Nidhoggr, the World-Worm (Dreaming Antiquity)

\subsubsection{Nidhoggr, the World-Worm (Dreaming Antiquity)}
\textit{Lore.} Beneath stone and sleep lies the slow memory of the world. Nidhoggr turns in aeons, dreaming of roads once walked and oaths once sworn.

\begin{quote}
Press your ear to the earth and wait. If it remembers you, it will answer.
\end{quote}

\paragraph{Glimpse the Ancient's Shadow (Low, 4 XP)} \emph{Action; Self; No.}\\
\textbf{Materials:} Pinch of dust from a worked stone.\\
\textbf{Effect:} +1 die to actions that identify, date, or interpret \emph{ancient} sites, scripts, or artifacts this scene; once this scene, ask one yes/no about the site's original purpose.\\
\textbf{Push It:} Add +1 Effect on one related roll, but suffer \emph{Fatigue 1}.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Drink from the Dreaming Deep (Low, 5 XP)} \emph{Instant; Self; No.}\\
\textbf{Materials:} Mouthful of clean water poured over stone, swallowed with eyes closed.\\
\textbf{Effect:} Learn one hidden factual detail about the immediate locale's \emph{past}. GM answers plainly or via a sensory echo.\\
\textbf{Cost:} Suffer \emph{Fatigue 1} and mark \emph{Exposure +1} as the dream clings.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Stone-Sleeper's Murmur (Standard, 7 XP)} \emph{Scene; Near (contact locus); No.}\\
\textbf{Materials:} Ear to bedrock, wall, or hewn pillar.\\
\textbf{Effect:} Once per beat while in contact, ask 1 question about a \emph{past event} that physically touched this stone; answers are fragmentary but truthful (max 3 questions/scene).\\
\textbf{Push It:} One answer is delivered with precise sensory clarity, but generate 1 SB (suit by GM).\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Awakened Chronicle (Standard, 9 XP)} \emph{Ritual (Significant Time); Zone; No.}\\
\textbf{Materials:} Chalk spiral and four touchstones from the site.\\
\textbf{Effect:} The zone ``replays'' a notable past moment as ghostly echoes all can witness (no harm). Participants gain +2 dice on \emph{one} Investigate/Recall about that event this scene.\\
\textbf{Push It:} Add a second moment from a different era, but mark +1 Obligation.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Dive into the World-Worm's Dream (High, 12 XP)} \emph{Scene; Self; No.}\\
\textbf{Materials:} Lie upon bare earth within a drawn circle of stones.\\
\textbf{Effect:} Ask up to \textbf{3} factual questions about the \emph{distant past} or \emph{buried truth} of this place, people, or item. Answers arrive as lucid dream signs.\\
\textbf{Cost (choose one):} Suffer \emph{Fatigue 2 \& Exposure +1} \emph{or} gain +3 dice to one reality-warping cast this scene and generate 2 SB immediately.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\quad \emph{Obligation:} 7 segments.
