% patrons/raeyn.tex
% Fate’s Edge — Patron: Raéyn, Mistress of the Sea (Tides, Change, Sea-Travel)

\subsubsection{Raéyn, Mistress of the Sea (Tides, Change, Sea-Travel)}
\textit{Lore.} Raéyn is the restless sea and the promise of a changing tide. She favors those who read currents, bargain with weather, and carry news between shores.

\begin{quote}
Mark the tide, name your course, and trust the wave-road.
\end{quote}

\paragraph{Tidemark's Favor (Low, 4 XP)} \emph{Scene; Self; No.}\\
\textbf{Materials:} A knotted length of salt-twine brushed with seawater.\\
\textbf{Effect:} Treat slick, swaying, or water-slicked footing as stable for you this scene. Gain \textbf{+1 die} on balance, boarding, or shipboard movement.\\
\textbf{Push It:} Extend to one ally in Close for one beat, but generate \textbf{1 SB} (Spades: shifting deck/hazards).\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Whisper of Currents (Low, 5 XP)} \emph{Instant; Near; No.}\\
\textbf{Materials:} A shell held to the ear while facing the wind.\\
\textbf{Effect:} Learn the \emph{safest near-term route} across water or coastline (reefs, eddies, patrol patterns) or gain \textbf{+1 die} to navigation checks along coasts or rivers for this scene.\\
\textbf{Push It:} Also learn where the \emph{fastest} route diverges, but mark \textbf{Exposure +1} (you leave a telltale wake).\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Changing Tide (Standard, 7 XP)} \emph{Scene; Zone (water-adjacent); No.}\\
\textbf{Materials:} A handful of pebbles cast in a crescent.\\
\textbf{Effect:} Bias water level and currents in the zone. Those moving \emph{with} the tide gain \textbf{+1 die} to traverse/board/evade; those moving \emph{against} it suffer \textbf{−1 die}. Small craft test to hold position.\\
\textbf{Push It:} Create a brief surge or drawdown (one beat): open a ford \emph{or} swamp a skiff on a Miss; mark \textbf{+1 Obligation}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Blessing of the Wave-Road (Standard, 9 XP)} \emph{Scene; Route (sea-to-sea); No.}\\
\textbf{Materials:} Two sea-glass markers dropped overboard at start and end points.\\
\textbf{Effect:} Consecrate a \emph{wave-road} between two visible marks. Allies traveling along it gain \textbf{+2 dice} to traverse, evade, or carry at sea; designated pursuers suffer \textbf{−1 die} to intercept on that leg. One active wave-road at a time.\\
\textbf{Push It:} For one beat, extend the route’s favor to an adjacent leg (shoal, cove, channel); mark \textbf{+1 Obligation}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Storm-Queen’s Hand (High, 12 XP)} \emph{Scene; Zone (sea/shore/sky above); No.}\\
\textbf{Materials:} A vial of rainwater gathered at \emph{three crossings}.\\
\textbf{Effect:} Shape a bound storm-band over the zone. Choose \textbf{two} modes at cast; you may switch one mode once per scene:
\begin{itemize}
  \item \textbf{Propulsion:} Favorable wind/sea; your vessel gains +1 band of movement per beat (or \textbf{+1 Effect} to maneuvers).
  \item \textbf{Concealment:} Veil of rain and spray; ranged targeting is impaired; \textbf{−1 die} to hostile sighting/tracking.
  \item \textbf{Smite:} \texttt{[AREA]} wave or lightning lash once per beat as an action; treat as a strong hazard (GM sets Cap by fiction).
\end{itemize}
\textbf{Push It:} Add a third mode for one beat, then the storm “slips” and the GM immediately spends \textbf{1 SB} on collateral; mark \textbf{+1 Obligation}.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\quad \emph{Obligation:} 7 segments.

\paragraph{Rivalry: Khemesh.} Raéyn’s currents \emph{counter} Khemesh’s abyssal \emph{pressure}. In scenes dominated by \emph{open water and weather}, Raéyn rites that shift tides/currents gain favorable fiction against \emph{Depth/Pressure} effects; in \emph{deep or confined} waters, Khemesh’s rites exert advantage. Adjudicate as Position/Effect shifts or modest DV nudges; SB remains the primary prompt.
