% --- Patron: The Traveler (Ways & Roads) ---
\subsubsection{The Traveler (Ways \& Roads)}
\textit{Lore.} Crossroads remember every footfall. The Traveler minds the stories that move between places.

\begin{quote}
Put one foot in a promise, and the road will meet you halfway.
\end{quote}

\paragraph{Road-Sense (Low, 4 XP)} \emph{Action; Self; Yes (navigation).}
\textbf{Materials:} A road-nail or waystone pebble.\\
\textbf{Effect:} Unerringly pick the fastest \emph{safe} route in Near/Far; \textbf{+1 die} to avoid ambushes and delays this leg/scene.\\
\textbf{Push It:} Also spot one hidden bypass; taking it creates \textbf{1 SB (Clubs)} elsewhere on the map.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Traveler’s Boon (Low, 5 XP)} \emph{Scene; Self/Ally; No.}
\textbf{Materials:} Tie thread around a wrist.\\
\textbf{Effect:} Ignore one level of difficult terrain or bureaucracy for this scene; \textbf{+1 effect} to overland progress/escape checks.\\
\textbf{Push It:} Extend to one additional ally; mark \textbf{1 SB (Diamonds)} as the road exacts a toll (favors, papers, attention).\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Waymark \textnormal{[PASSAGE]} (Standard, 7 XP)} \emph{Action; Near; No.}
\textbf{Materials:} Chalk mark at eye level.\\
\textbf{Effect:} Declare a lane as permitted/easy: allies on that lane gain better flow (Position/Effect bump or ignore one obstacle).\\
\textbf{Push It:} The lane persists between scenes until disturbed; first enemy who exploits it \emph{forces 1 SB (Spades)} on your party.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Bridge the Mile \textnormal{[TRANSPORT]} (Standard, 9 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} Two pinches of road-dust clapped together.\\
\textbf{Effect:} Relocate a willing target within Far along a visible or named route; arrivals are steady but noticed.\\
\textbf{Push It:} Carry one extra ally or a small bundle; arrivals are off-balance (worse Position for one beat).\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Crown of Crossings (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A brass compass missing its needle.\\
\textbf{Effect:} You call the Road: allies gain \textbf{+1 die} to move/evade; pursuit suffers \(-1\) die. Once, declare “the long way is short” to finish a travel clock segment for free.\\
\textbf{Push It:} Also seal a hostile route (like a temporary [WARD] against passage) for one beat; generates \textbf{2 SB (Clubs/Diamonds)} in border complications.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.
