% ============================================================
\section{Encounters}
% ============================================================

All encounters -- combat, social, environmental, magical -- use the same core engine: 
\textbf{Position}, \textbf{Effect}, \textbf{DV}, \textbf{Boons}, and \textbf{Story Beats (SB)}. 
Clocks track progress toward resolution.

\subsection{Common Framework}
\begin{itemize}[leftmargin=*]
  \item \textbf{State Intent \& Approach.} What do you want to achieve, and how?
  \item \textbf{Set Position \& DV.} GM determines risk (Dominant/Controlled/Desperate) and difficulty (2--7+).
  \item \textbf{Roll Attribute + Skill.} Each 6+ = success; each 1 = SB for the GM.
  \item \textbf{Apply Outcome.} Use the Outcome Matrix (Clean Success, Success \& Cost, Partial, Miss).
  \item \textbf{Spend SB.} GM introduces complications, shifts Position, or advances clocks.
\end{itemize}

\subsection{Clocks in Encounters}
Clocks track cumulative progress or increasing pressure:
\begin{itemize}[leftmargin=*]
  \item \textbf{Progress Clock (4--8 segments):} Fill to succeed (e.g., persuasion, escaping a maze).
  \item \textbf{Threat Clock (4--8 segments):} Fill to trigger a disaster (e.g., reinforcements, structural collapse).
  \item \textbf{Contest Clocks:} Two clocks -- one for each side; first to fill wins.
\end{itemize}
Tick clocks on \textbf{Partial} (1 tick) or \textbf{Clean Success} (2 ticks). Misses may tick opposition clocks.

% ============================================================
\section{Combat Encounters}
% ============================================================

\subsection{Rhythm of Battle}
No fixed initiative. GM spotlights based on fiction. Each turn: \textbf{1 Action + 1 Move}.

\subsubsection*{Movement Basics}
\begin{itemize}[leftmargin=*]
  \item \textbf{Move:} Shift one range band (Close$\leftrightarrow$Near$\leftrightarrow$Far) or reposition within same band.
  \item \textbf{Dash (Action):} Move two bands; your next defense is \textit{Desperate}.
  \item \textbf{Disengage (Action):} Then your Move does not provoke opportunity attacks. You cannot Attack this turn.
\end{itemize}

\subsubsection*{Opportunity Attacks}
Leaving an enemy's \textbf{Close} range without \textit{Disengage} provokes one free melee attack (\textit{Controlled} Position, no action cost). Additionally, the following actions provoke an opportunity attack if performed while \textbf{engaged in Close} with an enemy:
\begin{itemize}[leftmargin=*]
  \item Making a ranged attack (bow, crossbow, thrown weapon).
  \item Casting a spell that does \textbf{not} have the \texttt{[TOUCH]} or \texttt{[WARD]} attribute.
\end{itemize}
The attacker makes one free melee attack at \textit{Controlled} Position.

\subsubsection*{Charge \& Skirmish (Full Turn Actions)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Charge:} Move two bands toward an enemy and make one melee attack. Costs 1 Fatigue. Next defense \textit{Desperate}. Requires open ground.
  \item \textbf{Skirmish:} Move two bands away from an enemy and make one ranged attack. Costs 1 Fatigue. Next defense \textit{Desperate}. Requires open ground.
\end{itemize}
Talents (e.g., \textit{Effortless Charge}, \textit{Relentless}) modify these.

\subsection{Health, Fatigue, and Harm}

\paragraph{Fatigue Track}
\begin{itemize}[leftmargin=*]
  \item \textbf{Track size = Body} (minimum 1).
  \item Mark Fatigue from exertion (Charge/Skirmish, forced marches, magic costs, Obligation overflow).
  \item \textbf{Effect of Fatigue:} Each point of Fatigue worsens your \textbf{Position} by one step (Dominant $\to$ Controlled $\to$ Desperate). If already \textit{Desperate}, each additional Fatigue instead imposes a \textbf{--1 die penalty} on all your rolls.
  \item Penalties from Fatigue and Harm \textbf{stack}.
\end{itemize}

\paragraph{Harm Levels}
\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Harm Level} & \textbf{Effect} \\
\midrule
Harm 1 & --1 die on actions directly related to the injury (GM's discretion). \\
Harm 2 & --1 die on \textbf{most} actions until treated. \\
Harm 3 & Incapacitated or dying. \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Taking Harm and Fatigue -- The Procedure}
When you are hit, follow these steps \textbf{in order}:

\begin{enumerate}[nosep]
  \item \textbf{Convert} -- Use the Armor Conversion Table to turn incoming Harm into Fatigue.  
        Minimum 1 Fatigue per hit, even if conversion would give 0.
  \item \textbf{Apply remaining Harm} --  
        \textit{First:} If any Harm remains, clear all existing Fatigue (the shock of a wound resets exhaustion).  
        \textit{Then:} Increase your Harm level by the number of Harm points remaining.
  \item \textbf{Apply Fatigue} from step 1. If this fills your Fatigue track, resolve Overflow (see below).
\end{enumerate}

\begin{table}[h]
\centering
\caption{Armor Conversion Table}
\begin{tabular}{lccc}
\toprule
\textbf{Armor Type} & \textbf{Incoming Harm 1} & \textbf{Incoming Harm 2} & \textbf{Incoming Harm 3} \\
\midrule
Light  & Fatigue 1             & Harm 1 + Fatigue 1    & Harm 2 + Fatigue 1 \\
Medium & Fatigue 1             & Fatigue 1             & Harm 2 + Fatigue 1 \\
Heavy  & Fatigue 1             & Fatigue 1             & Fatigue 2 \\
\bottomrule
\end{tabular}
\end{table}
\noindent The table shows the result of conversion. For example, Light Armor vs Harm 2 becomes “Harm 1 + Fatigue 1”. Minimum 1 Fatigue always applies.

\paragraph{Overflow (Fatigue Track Fills)}
\begin{itemize}[leftmargin=*]
  \item After adding Fatigue from step 3, if your Fatigue total equals or exceeds your Body (track size):
  \begin{enumerate}[nosep]
    \item Increase \textbf{Harm} by one level.
    \item \textbf{Clear all Fatigue}.
    \item Any excess Fatigue beyond the track size is lost.
  \end{enumerate}
  \item This can happen multiple times in one hit (e.g., with Body 1 or 2).
\end{itemize}

\paragraph{Examples}
\begin{itemize}[leftmargin=*]
  \item \textbf{Light Armor, Harm 2:} Convert 1 $\to$ 1 Fatigue, remaining Harm 1. Apply Harm (clears Fatigue, Harm 0$\to$1), then add 1 Fatigue. Result: Harm 1, Fatigue 1.
  \item \textbf{Heavy Armor, Harm 3:} Convert 3 $\to$ 2 Fatigue, remaining Harm 0. No Harm taken, add 2 Fatigue. Result: Fatigue +2.
  \item \textbf{Heavy Armor, Harm 1:} Convert 1 $\to$ 0 Fatigue, but minimum 1 Fatigue. No Harm. Result: Fatigue +1.
  \item \textbf{Overflow:} Body 2, currently 1 Fatigue, Heavy Armor hit with Harm 3. Convert 3 $\to$ 2 Fatigue. Add Fatigue: 1+2 = 3, track full. Overflow triggers: Harm 0$\to$1, clear Fatigue. Result: Harm 1, Fatigue 0.
\end{itemize}

\paragraph{Recovery}
\begin{itemize}[leftmargin=*]
  \item \textbf{Fatigue:} Short rest (1 hour) removes 2 Fatigue; full night's rest removes all.
  \item \textbf{Harm:} Requires medical care (Medicine roll) or magical healing; natural recovery takes a week per level.
\end{itemize}

\subsection{Weapons \& Armor (Quick Reference)}
\begin{center}
\begin{tabular}{lccc}
\toprule
\textbf{Weapon Type} & \textbf{Close} & \textbf{Near} & \textbf{Tempo} \\
\midrule
Light  & +2 dice & +1 dice & Fast (Move + Shoot) \\
Medium & +1 dice & +2 dice & Standard (Move or Shoot) \\
Heavy  & --1 dice & +3 dice & Slow (Load, then Shoot) \\
\bottomrule
\end{tabular}
\end{center}
Armor uses the conversion table above. Shields and weapon tags (Accurate, Brutal, etc.) are optional (see the full rules).

% ============================================================
\section{Social Encounters}
% ============================================================

Use \textbf{Persuasion Clocks} for extended negotiations, debates, or intimidation.

\subsection{Setting Up}
\begin{itemize}[leftmargin=*]
  \item \textbf{Target's Resolve:} Easy (4-segment), Moderate (6-segment), Hard (8-segment).
  \item \textbf{Initial Position:} Based on relationship, leverage, and setting (Dominant = mutual trust, Desperate = hostile audience).
  \item \textbf{DV:} Standard ladder (3--7+) with modifiers for cultural gaps, time pressure, etc.
\end{itemize}

\subsection{Non-Face Characters Can Help}
\begin{itemize}[leftmargin=*]
  \item \textbf{Setup:} Create an advantage -- produce evidence, cause a distraction, or perform a display of force. Roll an appropriate skill. On success, grant the primary speaker +1 Position or +1 die.
  \item \textbf{Assist:} Spend 1 Boon to give +1 die (narrated as a timely interjection or meaningful glance).
  \item \textbf{Interrupt:} On a Miss, a non-face character can take the fallout (e.g., absorb a social rebuke) by marking 1 Fatigue.
\end{itemize}

\subsection{Persuasion Rolls}
Each distinct argument or lever advances the clock:
\begin{itemize}[leftmargin=*]
  \item \textbf{Clean Success:} +2 segments.
  \item \textbf{Partial:} +1 segment; GM may spend SB for a complication (e.g., "They agree, but you'll owe a favor").
  \item \textbf{Miss:} No progress; GM may tick an \textbf{Opposition Clock} (4--6 segments).
\end{itemize}
When the Opposition Clock fills, the target hardens -- Position worsens or the ask becomes impossible.  
When the Persuasion Clock fills, the target agrees as framed.

\paragraph{Example: Persuading a Guard}
\begin{quote}
The party wants a guard to let them pass. The GM sets a 6-segment Persuasion Clock (Moderate resolve). Initial Position: Controlled (the guard is suspicious but not hostile). The rogue spends 1 Boon to \textit{Assist} with a whispered lie; the party’s face rolls Presence+Sway (DV 3) and gets a Partial: +1 segment, but the GM spends the generated SB to add a complication – the guard demands a bribe. The party pays (lose a minor asset use) and continues. Three successful arguments later, the clock fills; the guard steps aside.
\end{quote}

% ============================================================
\section{Environmental Encounters}
% ============================================================

\subsection{Terrain and Positioning}
\begin{itemize}[leftmargin=*]
  \item \textbf{Height Advantage:} +1 die or +1 Effect for attacks from above.
  \item \textbf{Cover:} Partial cover (+1 DV to target you); Full cover (cannot be targeted directly). You may \textbf{Hide} behind full cover (Stealth vs Perception).
  \item \textbf{Difficult Terrain:} Costs \textbf{2 Move actions} to traverse one band (or a full action to Dash). Charge/Skirmish require open ground unless the \textit{Relentless} talent is used.
\end{itemize}

\subsection{Hazards and Clocks}
Active hazards (fire, flooding, toxic gas) are represented by \textbf{Threat Clocks}:
\begin{itemize}[leftmargin=*]
  \item Tick the clock on missed rolls, GM SB spends, or at scene intervals.
  \item When the clock fills, the hazard triggers (e.g., collapse, wave of fire).
  \item Characters may attempt to \textbf{slow} or \textbf{disable} the hazard (e.g., Tinker to seal a gas leak, Magic to freeze spreading flames). Treat as a \textbf{Progress Clock}.
\end{itemize}
\textit{Example: Collapsing Bridge [6]} -- ticks on each Partial or Miss. At 3 segments, footing becomes Desperate; at 6, the bridge falls.

% ============================================================
\section{Magical Encounters (Brief)}
% ============================================================

\subsection{Adjudicating Spells in Combat}
\begin{itemize}[leftmargin=*]
  \item \textbf{Free Caster (Spellcraft):} Weave then Cast (2 actions). Push (mark Fatigue) to compress to 1 action. Casting while \textit{Engaged} worsens Position by one step unless the spell has \texttt{[TOUCH]} or \texttt{[WARD]}.
  \item \textbf{Runekeeper (Rites):} Invoke a Rite as 1 action; mark Obligation. \textit{Push It} (+1 Obligation) for increased effect.
  \item \textbf{Invoker (Symbols):} Ritual takes DV+1 rounds. \textit{Crack the Seal} (instant) marks +2 Obligation (or +3 for High Rites), Symbol becomes Compromised.
  \item \textbf{Cantor (Songs):} Sing a Song (1 action, resolves at start of next turn). Push to resolve immediately (mark Fatigue + advance Corruption).
\end{itemize}

\subsection{Traps and Wards (Quick Look-up)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Discover:} Notice vs DV 4--6.
  \item \textbf{Disable:} Contested roll vs trap's tick. Clean Success = neutralized; Partial = disable but generate SB; Miss = trap activates.
  \item \textbf{Activation:} GM spends SB to trigger trap (damage, condition, clock advance).
  \item \textbf{Wards:} Magical barriers with a DV (3--6). Crossing without dispelling triggers the effect and ticks any relevant \textbf{Leash} by the ward's DV.
\end{itemize}

\subsection{Summoning and Spirits (Quick Look-up)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Leash Capacity:} Cap + Spirit segments.
  \item \textbf{Ticks:} Spirit takes Harm (1 per Harm level); acts against nature (1); crosses a ward without dispelling (ticks = ward's DV).
  \item \textbf{Leash fills:} Spirit acts once to its nature (potentially hostile) then departs.
  \item \textbf{Boon Finesse:} Summoner may spend 1 Boon to clear 1 tick (once per round).
\end{itemize}

\subsection{Extended Rituals and Downtime Encounters}
Use a \textbf{Progress Clock} (4--12 segments). Each segment represents a scene or a block of time. Players describe their approach and roll an appropriate skill; success advances the clock. Complications (GM SB) may add segments, require rare components, or draw unwanted attention.

% ============================================================
\section{Quick Reference Table}
% ============================================================
\begin{center}
\begin{longtable}{p{2.2cm}p{2.2cm}p{2.2cm}p{2.2cm}}
\toprule
\textbf{Encounter} & \textbf{DV (Default)} & \textbf{Typical Skills} & \textbf{Common Clocks} \\
\midrule
Combat & 3 (Controlled) & Melee, Ranged, Athletics, Command & Mob Overwhelm [6], Hazard [4-8] \\
Social  & 3 (Controlled) & Sway, Subterfuge, Insight, Command & Persuasion [4--8], Opposition [4] \\
Environmental & 3--4 (varies) & Athletics, Survival, Tinker, Stealth & Collapse [6], Flood [6], Chase [4] \\
Magical & 4--6 (specific) & Arcana, Lore, Spirit+Resolve & Leash [Cap+Spirit], Ritual [6--12] \\
\bottomrule
\end{longtable}
\end{center}

\subsection*{GM Reminders}
\begin{itemize}[leftmargin=*]
  \item \textbf{Let non-specialists contribute} -- use Setup and Assist actions, and allow creative skill use.
  \item \textbf{Spend SB to escalate, not punish.} A well-spent SB changes the situation, creating new choices.
  \item \textbf{Clocks make progress visible.} When a clock ticks, describe what changes in the fiction.
  \item \textbf{Armor \& Fatigue:} Follow the step-by-step procedure: convert, apply remaining Harm (clears existing Fatigue), then add Fatigue. Minimum 1 Fatigue per hit.
\end{itemize}