% --- Fate's Edge SRD --- Section 4: Talents & Equipment (Revised) ---
\section{Skills, Talents, and Equipment}

\subsection{Core Skills (Condensed)}
An action roll pairs an \textbf{Attribute} with a \textbf{Skill}. The GM sets Position and DV from the fiction. Core skills include:

\begin{itemize}[leftmargin=*]
  \item \textbf{Combat:} Melee, Ranged, Brawl, Tactics.
  \item \textbf{Physical:} Athletics, Stealth, Endurance, Craft, Survival.
  \item \textbf{Social:} Sway, Command, Deception, Performance, Insight.
  \item \textbf{Knowledge:} Lore, Investigation, Medicine.
  \item \textbf{Magic:} Arcana, Ritual.
\end{itemize}

\begin{multicols}{2}
\begin{itemize}[leftmargin=*]
  \item \textbf{Arcana} – magic theory, wards, free casting, counter-spelling.
  \item \textbf{Athletics} – running, climbing, jumping, swimming.
  \item \textbf{Brawl} – unarmed combat, grappling, improvised weapons.
  \item \textbf{Command} – leadership, intimidation, issuing orders.
  \item \textbf{Craft} – building, repairing, creating tools or structures.
  \item \textbf{Deception} – lying, bluffing, misdirection, forgery.
  \item \textbf{Endurance} – resisting fatigue, poison, harsh weather.
  \item \textbf{Insight} – reading emotions, detecting lies, understanding motives.
  \item \textbf{Investigation} – research, deduction, analysis of clues.
  \item \textbf{Lore} – history, culture, religion, customs, local knowledge.
  \item \textbf{Medicine} – healing, anatomy, first aid, surgery.
  \item \textbf{Melee} – swords, axes, close-quarters weapons.
  \item \textbf{Performance} – singing, acting, oration, musical display.
  \item \textbf{Ranged} – bows, crossbows, thrown weapons.
  \item \textbf{Ritual} – performing rites, invocations, exorcisms, consecrations.
  \item \textbf{Stealth} – moving unseen, hiding, camouflage.
  \item \textbf{Survival} – tracking, foraging, predicting weather, navigating wilderness, animal handling.
  \item \textbf{Sway} – persuasion, negotiation, charm.
  \item \textbf{Tactics} – battlefield strategy, unit coordination, ambush planning.
\end{itemize}
\end{multicols}

\noindent
\textit{Note:} \textbf{Survival} replaces \textit{Nature}, covering all wilderness and animal interactions. \textbf{Insight} replaces \textit{Notice} for social perception; use \textbf{Investigation} for physical clue-finding. \textbf{Ritual} is for actively performing magic (Rites, Songs, invocations); \textbf{Arcana} covers theory and identification.  
Optional modules (e.g., \textit{Psionics}) may add additional skills.

\noindent
\textbf{Common uses:} Use \textbf{Craft} for mechanical locks and traps; \textbf{Arcana} or \textbf{Lore} for magical wards. \textbf{Sway} gets you past people; \textbf{Stealth} and \textbf{Craft} get you through mechanisms.

\subsection{Talents}

Talents are the building blocks of character specialization. Each costs XP. \textbf{Passive} talents are always on; \textbf{active} talents require an action or reaction and cannot be used simultaneously unless noted. This list covers the core archetypes; many more exist in the full guides.

\paragraph{Talent Costs}
\begin{itemize}
  \item 2–3 XP: Minor edge (small bonus, once per scene).
  \item 4–6 XP: Major edge (permanent +1 die in a niche, access to magic).
  \item 7–10 XP: Prestige (campaign‑defining).
  \item 11+ XP: Epic (legendary, story‑shaping).
\end{itemize}

\subsubsection*{Magic Access Talents}
\begin{itemize}
  \item \textbf{Spellcraft (6 XP)} – enables freeform casting.
  \item \textbf{Familiar (2 XP)} – required for Patron's Gift, binds a Thiasos.
  \item \textbf{Codex (4 XP)} – needed for Runekeeper Rites and Obligation.
  \item \textbf{Patron's Symbol (4 XP)} – allows Invoker rituals.
  \item \textbf{Cantor's Path (8 XP)} – allows Songs (Corruption track).
  \item \textbf{Pact-Whisperer (2 XP)} – base summoning access.
\end{itemize}

\subsubsection*{Striker (Melee/Ranged Damage)}
\begin{itemize}
  \item \textbf{Weapon Master (5 XP)}: +2 dice with a chosen weapon; once/scene +1 Effect with that weapon. \textit{Passive + active.}
  \item \textbf{Flurry Strike (6 XP)}: Make two attacks as one action (each –1 die). \textit{Active, once/scene.}
  \item \textbf{Duelist's Edge (6 XP)}: +1 die when engaged with a single opponent; ignore the first Harm 1 or 2 from that opponent once/scene. \textit{Passive + active.}
  \item \textbf{Backstab (6 XP)}: When attacking from stealth, deal +1 Harm and ignore 1 point of Armor. \textit{Active, once/scene.}
  \item \textbf{Shadow Dance (10 XP)}: After a successful Backstab, test Stealth to re-enter stealth; on success, clear 1 Fatigue or improve Position by 1. \textit{Active, once/scene.}
  \item \textbf{Effortless Charge (4 XP)}: Ignore the Fatigue cost of Charge once/scene.
  \item \textbf{Effortless Skirmish (4 XP)}: Ignore the Fatigue cost of Skirmish once/scene.
  \item \textbf{Relentless (6 XP)}: May Charge or Skirmish over difficult terrain (Fatigue cost still applies). \textit{Passive.}
\end{itemize}

\subsubsection*{Defender (Tank, Protection)}
\begin{itemize}
  \item \textbf{Defensive Survival (3 XP)}: +1 die to defense rolls while engaged in melee; once/scene convert Harm 1 from melee to Fatigue. \textit{Passive + active.}
  \item \textbf{Conditioning (4 XP)}: Body counts as +1 for Fatigue track; +1 die to resist Fatigue overflow effects. \textit{Passive.}
  \item \textbf{Tactical Movement (4 XP)}: May move within engagement range as a Move action (instead of full action); once/scene disengage from Close as a Move action. \textit{Passive + active.}
  \item \textbf{Protect (action)}: Any character can use the \textit{Protect} action (see Action Options); no talent required.
  \item \textbf{Stubborn Pursuer (6 XP)}: When an opponent within Close range Disengages, you may use your reaction to move with them, staying in Close range. \textit{Reactive.}
\end{itemize}

\subsubsection*{Controller (Area, Debuff, Terrain)}
\begin{itemize}
  \item \textbf{Battlefield Mastery (6 XP)}: When engaged with 3+ opponents, for 3 exchanges: melee attacks gain +1 Effect, enemies act at –1 die, your Position improves by one step, and you may convert one Harm 1 per exchange to Fatigue. \textit{Active, once/scene.}
  \item \textbf{Spellcraft (6 XP)}: Enables freeform spellcasting (see Magic section). Use [TAGS] to create area effects, binds, walls, etc.
  \item \textbf{Ritual Mastery (4 XP)}: Reduce casting time for major rituals by half; GM banks 1 SB on any ritual backlash. \textit{Passive.}
\end{itemize}

\subsubsection*{Support (Healing, Buffs, Teamwork)}
\begin{itemize}
  \item \textbf{Field Dressing (2 XP)}: When an ally would take Harm that would incapacitate them, you may use your reaction to stabilize them; they remain at Harm 2 instead of 3 for the scene. \textit{Active, once/scene.}
  \item \textbf{Medic's Hand (3 XP)}: When you stabilize an ally, also clear 1 Fatigue or downgrade Harm 2→1. Spend 1 Boon to do both. \textit{Reactive.}
  \item \textbf{Inspire (3 XP)}: Once/scene, spend 1 Boon and choose: a bonded ally gains +1 Boon and +1 die on their next roll; or all allies in Near gain +1 die on their next roll. \textit{Active.}
  \item \textbf{Setup (action)}: Any character can use the \textit{Setup} action (see Action Options); no talent required.
  \item \textbf{Command Presence (4 XP)}: +1 die on leadership and intimidation rolls. \textit{Passive.}
\end{itemize}

\subsubsection*{Scout (Stealth, Infiltration, Awareness)}
\begin{itemize}
  \item \textbf{Keen Senses (2 XP)}: +1 die on perception checks to spot danger or hidden details. \textit{Passive.}
  \item \textbf{Trackless Step (2 XP)}: Once/scene, leave no trail for the remainder of the scene. \textit{Active.}
  \item \textbf{Urban Navigation (2 XP)}: Never get lost in cities. \textit{Passive.}
  \item \textbf{Silk Palms (2 XP)}: When you bribe or gift a minor official, gain a temporary \textit{String: Bought Silence}. Activate it to reduce DV by 1 on a Petition or Broker action involving that official (consumed). \textit{Active.}
  \item \textbf{Camouflage (4 XP)}: In natural terrain, spend an action to blend in; gain +2 dice to Stealth while remaining still, pursuers suffer –1 die to detect you. \textit{Active, once/scene.}
\end{itemize}

\subsubsection*{Leader (Tactics, Morale, Coordination)}
\begin{itemize}
  \item \textbf{Command Presence (4 XP)}: +1 die on leadership and intimidation rolls. \textit{Passive.}
  \item \textbf{Network Weaver (4 XP)}: Gain a minor contact in each new settlement you visit. \textit{Passive.}
  \item \textbf{Inspire (3 XP)}: As above.
  \item \textbf{Tactical Relay (3 XP)}: Once/scene, spend 1 Boon; all allies currently acting in Near gain +1 Position on their next action. \textit{Active.}
\end{itemize}

\subsubsection*{Monk (Unarmed Specialist)}
\begin{itemize}
  \item \textbf{Disciplined Body (3 XP)}: +1 die to unarmed attacks; once/scene convert Harm 1→Fatigue; once/scene improve Position by one step. \textit{Passive + active.}
  \item \textbf{Open Hand Technique (2 XP)}: On a successful unarmed attack, choose one: knock target prone, push them back one band, or force them to drop a held item. \textit{Active, once/scene.}
  \item \textbf{Iron Fist Way (6 XP)}: +1 die to unarmed attacks; strikes count as enchanted for harming supernatural creatures. \textit{Passive.}
  \item \textbf{Flowing Spirit Way (8 XP)}: Convert 1 Harm per attack to Fatigue; +1 die against fear or charm effects. \textit{Passive + active.}
  \item \textbf{Perfect Timing Way (7 XP)}: Twice per scene, improve Position by +1; +1 die to reaction rolls. \textit{Active.}
\end{itemize}

\subsection{Equipment}

\subsubsection{Weapons by Weight Class}
\begin{itemize}
  \item \textbf{Light (4 XP)}: Fast, concealable.
  \item \textbf{Medium (8 XP)}: Balanced.
  \item \textbf{Heavy (12 XP)}: Slow, powerful.
\end{itemize}

\paragraph{Melee Modifiers}
\begin{center}
\begin{tabular}{llll}
\toprule
Weight & Close & Near & Notes \\
\midrule
Light  & +2d & +1d & Quick \\
Medium & +1d & +2d & Set 1/scene or –1d first attack \\
Heavy  & –1d & +3d & Set 1/scene or –2d first attack \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Ranged \& Tempo}
\begin{center}
\begin{tabular}{lllll}
\toprule
Weight & Tempo & Close & Near & Far \\
\midrule
Light  & Fast   & Controlled & +1d & --- \\
Medium & Standard & Desperate & +2d & +1d \\
Heavy  & Slow    & Desperate & +1d & +3d \\
\bottomrule
\end{tabular}
\end{center}
\textit{Tempo: Fast = Move+Shoot; Standard = Move or Shoot; Slow = Set/Brace, no Move+Shoot.}

\subsubsection{Armor Conversion}
\begin{center}
\begin{tabular}{llll}
\toprule
Armor & XP & Conversion & Penalty \\
\midrule
Light  & 4  & 1 Harm→1 Fatigue & --- \\
Medium & 8  & 2 Harm→1 Fatigue & –1d physical \\
Heavy  & 12 & 3 Harm→2 Fatigue & –2d physical, no sprint \\
\bottomrule
\end{tabular}
\end{center}
Conversion applies \textit{per Harm instance} before marking Fatigue. Minimum 1 Fatigue per hit.

\subsubsection{Shields (Optional)}
\begin{center}
\begin{tabular}{llll}
\toprule
Shield & XP & Benefit & Tradeoff \\
\midrule
Buckler & 4 & +1d Defend or +1 DV once & Off-hand \\
Heater  & 8 & +1d Defend; 1 Harm→Fatigue & –1d Ranged \\
Pavise  & 12 & Plant: heavy cover cone & Bulky, slow \\
\bottomrule
\end{tabular}
\end{center}
Using a shield restricts weapon size to Light or Medium.

\subsection{Equipment Condition \& Upkeep}

All gear can become \textbf{Neglected} or \textbf{Compromised}:
\begin{itemize}
  \item \textbf{Neglected}: Weapons –1 die; Armor conversion worsened by 1 step.
  \item \textbf{Compromised}: Weapons –1 die first attack/round; Armor provides no conversion.
\end{itemize}
\textbf{Repair:}
\begin{itemize}
  \item Short rest + tools: Remove Neglected.
  \item A scene of work or a smith: Remove Compromised.
  \item Enchanted items require the same repairs; enchantments are inactive while Compromised.
\end{itemize}

\subsection{Magical Items (Enchantments)}

Enchantments are permanent magical modifications treated as Talents (XP cost). They follow the same stacking rules: only one active enchantment per item unless noted. Enchanted items count as Assets for upkeep.

\subsubsection*{Sample Enchantments}
\begin{itemize}
  \item \textbf{Keen Edge (2 XP)}: +1 die vs armored targets.
  \item \textbf{Flaming Blade (4 XP)}: +1 fire damage, +1 Effect vs cold creatures.
  \item \textbf{Soulfire Weapon (6 XP)}: Ignores 1 point of armor/resistance.
  \item \textbf{Shadowweave Armor (2 XP)}: +1 die to Stealth while moving.
  \item \textbf{Runed Plate (4 XP)}: Reduce magical Backlash by 1 SB.
  \item \textbf{Boots of Sure Step (3 XP)}: Ignore difficult terrain.
\end{itemize}

\subsubsection*{Enchantment Limits}
\begin{itemize}
  \item Max enchantments = item's Tier (Light=1, Medium=2, Heavy=3).
  \item Identical benefits do not stack.
  \item Removing an enchantment: ritual (Wits+Arcana DV 4) + 1 XP per enchantment level.
\end{itemize}

\subsection{Worked Example}
\emph{Sable earns 5 XP from a Drive and trades 2 Boons for 1 XP. She buys \textbf{Spellcraft} (6 XP). Later, she finds a \textbf{Keen Edge} longsword (2 XP enchantment) – she pays the XP and now has +1 die against armored foes.}