% --- Fate's Edge SRD --- Section 18: Campaign Frame: The Crown Spread (Optional) ---
\section{Campaign Frame: The Crown Spread (Optional)}
\label{sec:crown-spread}

The \textbf{Crown Spread} is a campaign‑framing tool that uses a spread of cards to establish the long arc of a story. It provides seeds for GMs and players to weave motifs, omens, and foreshadowed events. Its use is entirely optional.

\subsection{Setup}
\begin{itemize}
  \item In Session 0 (or before the campaign begins), lay out 5--7 cards in a semicircle (the "Crown"). Use either the \textbf{Deck of Consequences} (see \S\ref{sec:deck-consequences}) or a standard card deck.
  \item Each card anchors a motif, omen, or looming event.
  \item Record the spread openly on a Campaign Sheet or digital log.
\end{itemize}

\subsection{Interpreting the Spread}
\begin{description}[leftmargin=1.5em, style=nextline]
  \item[Position 1 (Root):] The underlying tension or theme of the campaign.
  \item[Position 2 (Crest):] A key faction or patron influence that will rise.
  \item[Position 3 (Crown):] The climax image or major confrontation the arc builds toward.
  \item[Position 4 (Left Hand):] A bond, ally, or relationship that anchors play.
  \item[Position 5 (Right Hand):] A rival, betrayer, or challenger who pressures the party.
  \item[Optional Positions 6 \& 7:] Expansions for setting‑wide twists (environmental, mystical, or political).
\end{description}

\subsection{Using the Spread in Play}
\begin{itemize}
  \item Each drawn card becomes a \textbf{Foreshadow Clock [4]} attached to its motif. Advance the clock when events lean toward that omen (e.g., when the GM spends SB on related complications, when a major scene involves that faction, or when a travel leg passes through an area tied to the motif).
  \item When a Foreshadow Clock fills, the motif manifests concretely in play -- a faction rises, a betrayer reveals themselves, a curse activates.
  \item Tie Spread cards to \textbf{Campaign Clocks} for pacing (see below).
\end{itemize}

\subsection{Campaign Clock}
The \textbf{Campaign Clock} tracks rising stakes across the entire arc.
\begin{itemize}
  \item Default size: [8] segments.
  \item Advance the Campaign Clock when: multiple SB overflows in a session (e.g., three or more 1s on a single roll), when travel legs resolve with major cost (e.g., a character reaches Harm 2), or when Spread omens manifest concretely.
  \item When the Campaign Clock fills, the Crown confrontation arrives. Play through its fallout as the campaign climax.
\end{itemize}

\subsection{Ending \& Legacy}
\begin{itemize}
  \item After the Crown confrontation resolves, hold an epilogue session.
  \item Resolve any remaining Foreshadow Clocks as epilogue vignettes.
  \item Players may mark \textbf{Legacy Bonds} -- new anchors for future campaigns or descendants.
\end{itemize}

\subsection{GM Quick Cues}
\begin{itemize}
  \item The Spread is not a railroad -- it foreshadows, not dictates.
  \item Reinterpret cards liberally as play evolves; symbols matter more than literal events.
  \item Remind players of their omens between arcs to build tension and payoff.
\end{itemize}