% --- Fate's Edge SRD --- Section 19: Tags & Outcomes Cheat Sheet ---
\section{Tags \& Outcomes Cheat Sheet}
\label{sec:tags-cheatsheet}

This section condenses common tags, outcomes, and modifiers into a one‑glance reference. It supplements the full explanations in \S\ref{sec:tags}.

\subsection{Core Outcome Matrix}
\begin{center}
\begin{longtable}{|p{2.8cm}|p{3.5cm}|p{8.2cm}|}
\hline
\textbf{Result} & \textbf{Meaning} & \textbf{GM/Player Guidance} \\
\hline
Successes $\ge$ DV, SB=0 & Clean Success & Intent achieved, no complication. \\
\hline
Successes $\ge$ DV, SB$\ge$1 & Success \& Cost & Intent achieved; GM spends SB for complication. \\
\hline
$0 <$ Successes $<$ DV & Partial & Progress with complication; player gains 1 Boon. \\
\hline
Successes $=$ 0 & Miss & No progress; GM escalates with SB; player gains 2 Boons. \\
\hline
\end{longtable}
\end{center}

\subsection{Common Tags (Abridged)}
\begin{center}
\begin{longtable}{|p{3.5cm}|p{10.0cm}|}
\hline
\textbf{Tag} & \textbf{Quick Reference Effect} \\ \hline
\texttt{[AREA]} & Affects all in a zone; collateral risk. \\ \hline
\texttt{[BANISH]} & Drive an Outsider toward departure; DV = target Cap. \\ \hline
\texttt{[BARRIER]} & Creates cover/obstruction with Integrity Clock. \\ \hline
\texttt{[BIND]} & Immobilizes target until broken. \\ \hline
\texttt{[CLEANSE]} & Remove or suppress a Condition (poison, curse, etc.). \\ \hline
\texttt{[COMMAND]} & Issue a clear order; resolve as social contest. \\ \hline
\texttt{[CONJURE]} & Create a useful object, cover, or hazard. \\ \hline
\texttt{[COUNTER]} & Interrupt a spell or rite in progress. \\ \hline
\texttt{[CURSE]} & Inflict a sticky hindrance with a clear release condition. \\ \hline
\texttt{[DISPEL]} & End or suppress an ongoing magical effect. \\ \hline
\texttt{[FORTIFY]} & Harden a target against a specific vector (fire, fear, etc.). \\ \hline
\texttt{[MARK]} & Tag a target; once/scene +1 die against it. \\ \hline
\texttt{[OATH]} & Bind parties to terms; breaking has teeth. \\ \hline
\texttt{[PASSAGE]} & Declare a route as permitted or easy (improves travel). \\ \hline
\texttt{[REVEAL]} & Expose illusions, disguises, or hidden clauses. \\ \hline
\texttt{[SANCTIFY]} & Consecrate a zone to a code or Patron. \\ \hline
\texttt{[SEAL]/[UNSEAL]} & Magically lock or unlock a container / door. \\ \hline
\texttt{[TRANSPORT]} & Move a target across an obstacle (blink, lift, pull). \\ \hline
\texttt{[UNWARD]} & Unmake or suppress a \texttt{[WARD]}. \\ \hline
\texttt{[VEIL]} & Obscure a person, object, or zone (--1 die to Notice). \\ \hline
\texttt{[WARD]} & Blocks specific entities (DV = target Cap to cross). \\ \hline
\end{longtable}
\end{center}

\subsection{Position \& Effect Shorthands}
\begin{itemize}
  \item \textbf{Position:} \emph{Dominant} (safe), \emph{Controlled} (standard), \emph{Desperate} (high stakes).
  \item \textbf{Effect:} \emph{Limited} (weak), \emph{Standard} (expected), \emph{Great} (strong, extra reach).
  \item GM can spend 1 SB to worsen Position one step; player can spend 1 Boon to improve Position one step.
\end{itemize}

\subsection{Backlash Reminders}
\begin{itemize}
  \item On Partial/Miss with Weave \& Cast (Free Casting), expect backlash colored by the dominant element (fire burns, fate twists, etc.).
  \item Backlash does not add SB directly -- it is a GM SB spend colored by the roll's element.
\end{itemize}

\subsection{Obligation Reminders}
\begin{itemize}
  \item Runekeeper Rite: Mark listed Obligation (usually +1); +1 more if Pushed.
  \item Invoker Ritual (via Symbol): Always +1 Obligation, even out of scene.
  \item Crack the Seal (instant cast): Mark +2 Obligation (+3 for High Rites); Symbol becomes Compromised.
  \item Multi‑symbol overload: carrying 4+ different Symbols adds +1 Obligation to the first Invoker ritual each scene.
\end{itemize}

\subsection{Magic Access Talent Costs (Quick Ref)}
\begin{itemize}
  \item \textbf{Spellcraft (6 XP):} Required for Weave \& Cast freeform magic.
  \item \textbf{Familiar / Thiasos (2 XP):} Required for Patron's Gift.
  \item \textbf{Codex (4 XP):} Full Runekeeper Rite access; Obligation applies.
  \item \textbf{Patron's Symbol (4 XP):} Invoker ritual access (Minor Asset).
\end{itemize}

\subsection{GM Quick Cues}
\begin{itemize}
  \item Default to \textbf{Near}, \textbf{Controlled}, \textbf{Standard Effect} unless fiction says otherwise.
  \item Use tags as handles for rulings -- if a move fits the tag, apply its shorthand.
  \item Complications escalate campaigns; use Clocks and the Deck of Consequences for follow‑through.
\end{itemize}