\section{New Core Talents}

These talents are available to any Runic Warrior regardless of Patron. They emphasize physical resilience, the conversion of Fatigue into advantage, and the bond between Runekeeper, Codex, and Thiasos.

\subsection{Foundation Talents (2-4 XP)}

\subsubsection{Flesh Codex (2 XP)}

\textbf{Requirement:} Body 3+

Your Rites are engraved on your body—tattoos, brands, or scarification. You have no physical Codex to steal, confiscate, or destroy.

\textbf{Effect:}
\begin{itemize}
\item Your Codex cannot be stolen or confiscated.
\item You always have access to your Rites (no action required to retrieve a Codex).
\item Damage to your skin (burns, deep wounds) may damage your Codex at the GM's discretion.
\end{itemize}

\subsubsection{Living Thiasos (3 XP)}

\textbf{Requirement:} A living Thiasos (warhorse, hound, wolf-ancestor, etc.)

Your Thiasos is not a pet. It is a partner.

\textbf{Effect:}
\begin{itemize}
\item Your Thiasos gains +1 Cap (maximum 5).
\item Once per scene, your Thiasos may act independently without requiring your action.
\item Your Thiasos is more resilient: when it would be reduced to 0 Harm, it remains conscious but Winded (-1 die to all actions) until healed.
\end{itemize}

\subsubsection{Oath-Bound (4 XP)}

\textbf{Requirement:} You have sworn a formal oath before a witness.

Your word is your bond. Your bond is your strength.

\textbf{Effect:}
\begin{itemize}
\item Once per session, when you act in direct service of a sworn oath, you may reroll a failed Body roll.
\item If you break your sworn oath, you lose access to this talent until you perform a witnessed atonement.
\end{itemize}

\subsubsection{Scarred Resilience (3 XP)}

\textbf{Requirement:} Body 2+

You have learned to read your own pain.

\textbf{Effect:}
\begin{itemize}
\item Once per scene, you may ignore 1 Fatigue from physical exertion (running, climbing, fighting, or invoking a Rite that costs Fatigue).
\item You may use this talent after the Fatigue is marked.
\end{itemize}

\subsection{Advanced Talents (6-8 XP)}

\subsubsection{Engraved Defense (6 XP)}

\textbf{Requirement:} Your Codex is a physical object (gorget, shield, staff, or sentient blade)

Your Codex is not just a record. It is armor.

\textbf{Effect:}
\begin{itemize}
\item Your Codex grants +1 Armor.
\item When you would take Harm, you may instead mark 1 Obligation and reduce the Harm by 1 (to a minimum of 0).
\item You cannot use this talent if your Codex is Compromised or destroyed.
\end{itemize}

\subsubsection{Shared Burden (6 XP)}

\textbf{Requirement:} A living Thiasos

You and your Thiasos share more than a covenant. You share weight.

\textbf{Effect:}
\begin{itemize}
\item When you mark Fatigue, you may transfer 1 point of it to your Thiasos.
\item Your Thiasos becomes Winded (-1 die to all actions) for the rest of the scene.
\item You may use this talent once per scene.
\end{itemize}

\subsubsection{Witnessed Scar (8 XP)}

\textbf{Requirement:} You have at least three permanent Obligation scars

Each scar is a payment made. Each payment is a memory.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may touch a scar and recall the event that caused it in perfect detail.
\item Gain +1 die on any roll directly tied to that event (tracking a foe you fought before, recalling a Patron's demand, navigating a place you bled in).
\end{itemize}

\subsubsection{Iron Resolve (7 XP)}

\textbf{Requirement:} Body 3+, Disciplined Body or similar resilience talent

You do not fall easily.

\textbf{Effect:}
\begin{itemize}
\item When you would be reduced to 0 Harm, you may instead remain standing with 1 Harm and mark 2 Fatigue.
\item You may use this talent once per scene.
\end{itemize}

\subsection{Prestige Talents (10-12 XP)}

\subsubsection{The Unbroken Vow (10 XP)}

\textbf{Requirement:} You have kept a sworn oath for at least one year (in-game time)

Some vows are not chains. They are bridges.

\textbf{Effect:}
\begin{itemize}
\item When you swear an oath before a witness, you gain +2 dice to all actions that serve that oath.
\item Breaking the oath causes you to immediately mark 2 Harm (Stress) and lose access to this talent until you perform a witnessed atonement.
\item You may have only one Unbroken Vow active at a time.
\end{itemize}

\subsubsection{Avatar of the Codex (12 XP)}

\textbf{Requirement:} Tier III+, your Codex is a physical object

You do not carry your Codex. You become it.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may become the living embodiment of your Codex for one scene.
\item Your Body increases by 2 (to a maximum of 6).
\item Your Thiasos fights beside you as a second combatant (taking its own actions).
\item At the end of the scene, you mark 2 Fatigue and 1 Obligation.
\end{itemize}

\subsubsection{The Scar That Remembers (11 XP)}

\textbf{Requirement:} Witnessed Scar, Tier III+

Your scars do not just record. They testify.

\textbf{Effect:}
\begin{itemize}
\item Once per session, you may touch a scar and speak a single sentence about the event that caused it.
\item That sentence is considered witnessed truth. In a hedge-court, it carries the weight of sworn testimony.
\item You cannot use this talent on the same scar twice.
\end{itemize}

\subsection{Patron-Specific Talents}

Each Patron family in this expansion includes unique talents for Runic Warriors. These are detailed in the Patron chapters.

Examples include:

\begin{itemize}
\item \textit{Dawn's Judgment} (Oath of Flame \& Light) — Smite undead and oath-breakers with radiant steel.
\item \textit{The Red Thread} (High Mother) — Store Fatigue as visible knots in your flesh.
\item \textit{Ghe'hai's Escape} (Lethai) — Once per session, vanish when cornered, reappearing in a safe location.
\item \textit{The Unseen Strike} (Ikasha) — First strike from concealment ignores 1 Armor.
\item \textit{Beast Form} (Savage Heart) — Partial transformation into a predator.
\end{itemize}

\subsection{Talent Synergies}

Some talents work well together:

\begin{itemize}
\item \textit{Flesh Codex} + \textit{Witnessed Scar} — Your body is both Codex and testimony. You cannot be disarmed.
\item \textit{Shared Burden} + \textit{Living Thiasos} — You and your Thiasos endure together. Neither falls alone.
\item \textit{Iron Resolve} + \textit{Avatar of the Codex} — You become nearly unkillable for one scene. The cost is high. The memory is a scar.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
