\section{Running Runic Warriors as NPCs}

Runic Warriors make compelling non-player characters because their covenant is visible. A scarred hand, an engraved gorget, a sentient blade that speaks unbidden—these are not subtle people. They are walking declarations of allegiance, and the world reacts to them accordingly.

This chapter provides guidance for using Runic Warriors as antagonists, allies, and neutral parties, followed by sample NPCs for each Patron family.

\subsection{Villains}

A Runic Warrior villain is not a schemer in a tower. They are warriors who have chosen the wrong Patron, or the right Patron for the wrong reasons.

\subsubsection{The Cursed Knight (Malachai)}

A knight who accepted Malachai's gift to survive a mortal wound. Now they cannot die—but neither can they stop killing. Their Codex is a silver coin that never spends. Their Thiasos is a Collector, a child with silver-wire eyes who watches in silence.

\textbf{Motivation:} Release from the curse, or the acquisition of enough power to force Malachai to bargain.

\textbf{Conflict:} The knight does not want to fight. They are tired. But the hunger drives them, and the Collector will not let them rest.

\textbf{Resolution:} The party can destroy the knight, cure them (requiring a journey to the Well of Ichor), or help them find a third path—a debt paid in memory rather than blood.

\subsubsection{The Rogue Bloodsworn (High Mother)}

A Daughter of the Cord who has broken her covenant. She still carries the Rites tattooed on her flesh, but she no longer pays the Cord its due. Now her scars weep. Her Hollowed bird has gone silent. She is dying by inches, and she will drag anyone she can with her.

\textbf{Motivation:} Find a way to sever the Cord without dying, or ensure that the Cord cannot use her debts against her family.

\textbf{Conflict:} She is desperate and unpredictable. She may help the party or betray them, depending on which advances her goal.

\textbf{Resolution:} The party can help her sever the Cord (a dangerous ritual requiring a witness and a sacrifice), turn her over to the Daughters, or kill her and scatter her ashes at a crossroads.

\subsubsection{The Fractured Knight (The Ninth)}

A warrior who glimpsed too much of the Ninth's truth. They now see patterns that do not exist, hear voices that predict the future, and move in ways that should be impossible. Their Codex is a broken mirror. Their Thiasos is a memory-spirit that speaks in overlapping voices.

\textbf{Motivation:} Find a way to unsee what they have seen, or embrace the fracture and become something beyond human.

\textbf{Conflict:} They are not evil. They are broken. Their actions serve a logic that no one else can perceive.

\textbf{Resolution:} The party can help them stabilize (requiring a ritual at a threshold and the sacrifice of a cherished memory), kill them, or leave them to their fractured path.

\subsection{Allies}

A Runic Warrior ally is a warrior who shares the party's goals, even if their methods differ. They are reliable in a fight, but their covenant may complicate things.

\subsubsection{The Dawn Warrior (Oath of Flame \& Light)}

A knight-commander of a border keep, sworn to protect the lowlands from incursions. Their Codex is an engraved gorget. Their Thiasos is a warhorse with milk-white eyes.

\textbf{Motivation:} Hold the line. Protect the innocent. Keep their oath.

\textbf{How They Help:} Combat support, tactical advice, sanctuary at their keep.

\textbf{Complication:} They cannot break their oath to the keep. If the party asks them to leave their post, they will refuse.

\subsubsection{The Ghe'hai Operative (Lethai Syncretic)}

A silk-blade assassin who serves the Lethai courts. Their Codex is a woven mask. Their Thiasos is a spider-spirit that weaves invisible threads.

\textbf{Motivation:} Complete their mission without leaving a trace.

\textbf{How They Help:} Infiltration, assassination, information gathering.

\textbf{Complication:} They answer to the Lethai courts. If the party's goals conflict with Lethai interests, the operative will vanish.

\subsubsection{The Beast Warrior (Savage Heart)}

A Sekogo hunter who wears the skin of a killed predator. Their Codex is a tooth. Their Thiasos is the spirit of the hunt.

\textbf{Motivation:} Hunt what must be hunted. Protect the wild from those who would tame it.

\textbf{How They Help:} Tracking, survival, combat against supernatural foes.

\textbf{Complication:} They do not understand civilization. Laws, debts, and social niceties are meaningless to them.

\subsection{Neutral Parties}

Some Runic Warriors are neither allies nor enemies. They have their own goals, and the party is incidental.

\subsubsection{The Itinerant Monk (The Traveler)}

A wanderer who carries no home and owes no allegiance. Their Codex is a waystone. Their Thiasos is a road-dog that walks beside them.

\textbf{Motivation:} Walk the road. See what is at the end.

\textbf{Neutral Stance:} They will help if the road leads the same direction. They will not risk their path for the party's convenience.

\subsubsection{The Cairn-Knight (Varnek Karn)}

A warrior who guards an ancestral burial ground. Their Codex is a rune-carved axe. Their Thiasos is an ancestral shade that fights beside them.

\textbf{Motivation:} Protect the cairn. Settle old wergild. Let the ancestors rest.

\textbf{Neutral Stance:} They will not leave the cairn. If the party needs their help, they must come to the cairn—and respect its laws.

\subsubsection{The Warden (The Sealed Gate)}

A sentinel who guards a threshold that must not be crossed. Their Codex is a shield. Their Thiasos is a standing stone that wakes when they speak.

\textbf{Motivation:} Keep the threshold sealed. Let nothing pass.

\textbf{Neutral Stance:} They do not care about the party's goals. They care only about the threshold. If the party tries to cross, they become enemies. If the party helps guard, they become allies.

\subsection{Sample NPCs by Patron Family}

\subsubsection{The Vowed}

\begin{itemize}
\item \textbf{Dawn Warrior:} \textit{Ser Roran of the Grey Gauntlet.} A knight who carries a silver gauntlet engraved with his oath. He never takes ransom. His warhorse is named Ledger. (Oath of Flame \& Light)
\item \textbf{Hospitaller:} \textit{Sister Bryn, retired.} A one-handed woman with a brass hook. She was a Daughter of the Cord once. Now she tends a hospice. Her Thiasos is a grey bell that rings when someone is dying. (Oath of Mercy \& Grace)
\item \textbf{Justicar:} \textit{Elodie Venn, Queen's Justicar.} A scarred woman with a silver gavel. Her Codex is a shield. Her Thiasos is a spirit-hound that tracks oath-breakers. (Mykkiel)
\item \textbf{Witch-Hunter:} \textit{Vigil-Commander Vorenus.} A Chain-Lantern prefect. His Codex is a censer. His Thiasos is a lantern that reveals hidden corruption. (Inquisitor Prime)
\item \textbf{Truth-Sworn:} \textit{The Archivist.} A silent woman who carries a silver mirror. Her Thiasos is a memory-spirit that records every word spoken in her presence. (The Witness)
\end{itemize}

\subsubsection{The Threshold}

\begin{itemize}
\item \textbf{Warden:} \textit{The Sentinel of the Silent Stair.} A figure in grey armor who stands at a chapel door. Their Codex is a shield. Their Thiasos is a standing stone that speaks in the voice of the dead. (The Sealed Gate)
\item \textbf{Itinerant Monk:} \textit{Brother Ash.} A wanderer who carries a waystone compass. His Thiasos is a road-dog with one blue eye and one brown. (The Traveler)
\item \textbf{Death Monk:} \textit{The Pallid Sister.} A grey-robed warrior who escorts the dying. Her Codex is a shepherd's crook. Her Thiasos is a spirit-shepherd that sings the names of the dead. (The Pale Shepherd)
\end{itemize}

\subsubsection{The Wild}

\begin{itemize}
\item \textbf{Root-Knight:} \textit{Elder Thorn.} A woman whose skin is bark. Her Codex is a living staff. Her Thiasos is a wolf-ancestor that speaks in growls. (Grimmir)
\item \textbf{Grave-Warden:} \textit{The Compost-Knight.} A figure in moss-covered armor. Their Codex is a bone-chime. Their Thiasos is a fungal familiar that speaks in spores. (The Carrion-King)
\item \textbf{Beast Warrior:} \textit{Mkara of the Spiral Throw.} A Sekogo hunter who wears a leopard pelt. Her Codex is a predator's tooth. Her Thiasos is the spirit of the hunt. (The Savage Heart)
\end{itemize}

\subsubsection{The Deep}

\begin{itemize}
\item \textbf{Eldritch Knight:} \textit{The Drowned Prince.} A figure in black glass armor. His Codex is a shard of abyssal glass. His Thiasos is a drowned spirit that speaks in bubbles. (Khemesh)
\item \textbf{Storm-Sworn:} \textit{The Thunder Shaman.} A Ykrul warrior who carries a drum made from storm-felled wood. Her Thiasos is a storm-spirit that crackles with lightning. (Xhak'Thul)
\item \textbf{Tide Warrior:} \textit{Captain Mira the Lucky.} A Sidhi pirate with a brass hook. Her Codex is a ship's wheel. Her Thiasos is a selkie that sings in storms. (Raéyn)
\end{itemize}

\subsubsection{The Shadow}

\begin{itemize}
\item \textbf{Shadow Warrior:} \textit{The Unseen.} A figure in a black feather cloak. Their Codex is a raven's quill. Their Thiasos is a raven-spirit that carries whispers. (Ikasha)
\item \textbf{Vowed Monk:} \textit{The Silent Blade.} A warrior who has sworn never to speak. Their Codex is a sealed-lips tattoo. Their Thiasos is a memory-spirit that speaks for them. (The Silent Choir)
\item \textbf{Ronin:} \textit{The Storm-Rider.} A masterless warrior on a grey horse. Their Codex is a storm-banner. Their Thiasos is a wind-spirit that rides beside them. (Aveh)
\end{itemize}

\subsubsection{The Web}

\begin{itemize}
\item \textbf{Web Warrior:} \textit{The Spider-Blade.} A Lethai operative in a silk mask. Their Codex is a spider-silk cord. Their Thiasos is a spider-spirit that weaves invisible threads. (Inaea)
\item \textbf{Serpent Warrior:} \textit{The Shed-Knight.} A warrior who has molted their identity. Their Codex is a shed fang. Their Thiasos is a serpent-spirit that speaks in sibilance. (Isoka)
\item \textbf{Cursed Warrior:} \textit{The Thorn-Sworn.} A figure with scars that weep. Their Codex is a silver coin that is always warm. Their Thiasos is a Collector. (Malachai)
\end{itemize}

\subsubsection{The Forge}

\begin{itemize}
\item \textbf{Artificer:} \textit{The Gear-Warrior.} A dwarf in a toolbelt. Their Codex is a schematic-slate. Their Thiasos is a clockwork construct that chitters and ticks. (Clockwork Monad)
\end{itemize}

\subsubsection{The Grave}

\begin{itemize}
\item \textbf{Cairn-Knight:} \textit{The Grave-Warden.} A warrior who guards a burial mound. Their Codex is a rune-carved axe. Their Thiasos is an ancestral shade. (Varnek Karn)
\item \textbf{Bloodsworn:} \textit{The Cord-Knight.} A Daughter of the Cord with Rites tattooed on her flesh. Her Codex is her skin. Her Thiasos is a hollowed bird with milk-white eyes. (The High Mother)
\end{itemize}

\subsubsection{The Fracture}

\begin{itemize}
\item \textbf{Fractal Warrior:} \textit{The Paradox-Sworn.} A warrior whose eyes see angles that should not exist. Their Codex is a broken mirror. Their Thiasos is a pattern-spirit. (The Ninth)
\item \textbf{Moon Knight:} \textit{The Silver-Sworn.} A warrior who fights only under moonlight. Their Codex is a silver disc. Their Thiasos is a shadow-cat. (Lunara)
\end{itemize}

\subsubsection{The Lethai Ghe'hai}

\begin{itemize}
\item \textbf{Spec Ops:} \textit{The Silk-Blade.} An elite operative with no name. Their Codex is a woven mask. Their Thiasos is a spider-spirit. They can kill a target in a crowded room and walk out without anyone knowing. (Lethai Syncretic)
\end{itemize}

\subsection{Running Runic Warriors in Combat}

Runic Warriors are front-line combatants. Use them as:

\begin{itemize}
\item \textbf{Tanks:} High Body, high Harm threshold, talents that convert damage to Fatigue or Obligation.
\item \textbf{Strikers:} Rites that enhance weapon damage, grant extra attacks, or ignore armor.
\item \textbf{Controllers:} Rites that bind foes, create difficult terrain, or force enemies to target them.
\end{itemize}

Remember that a Runic Warrior's Codex can be targeted. An enemy who sunders a gorget or shatters a staff has done more than damage equipment—they have severed the Runekeeper's connection to their Rites. Use this as a tactical option for intelligent foes.

\subsection{Runic Warriors in Faction Politics}

\begin{itemize}
\item \textbf{Chain-Lanterns:} Suspicious of Codexes that cannot be confiscated. Tattoos and brands are harder to burn than books.
\item \textbf{Daughters of the Cord:} View Bloodsworn as sisters or rivals. A Bloodsworn who serves the High Mother may be welcomed; one who serves another Patron may be watched.
\item \textbf{Aeler Vent-Priors:} Respect Artificers and Cairn-Knights. A Codex that is also a tool is efficient. Efficiency is virtue.
\item \textbf{Lethai Courts:} Ghe'hai operatives are trusted—until they are not. A Lethai Runic Warrior is a weapon, and weapons can be turned.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
