\section{Sample NPCs by Patron Family}

This chapter provides ready-to-use non-player characters for each Patron family. Each entry includes a brief description, notable traits, their Codex and Thiasos, a roleplaying hook, and suggested mechanical stats.

\subsection{The Vowed}

\subsubsection{Dawn Warrior — Ser Roran of the Grey Gauntlet}

A knight in battered plate armor, his left hand encased in a silver gauntlet engraved with a single word: \textit{Paid}. He rides a grey warhorse named Ledger. His face is scarred, his eyes tired, his voice quiet.

\textbf{Codex:} The Grey Gauntlet. Engraved with Rites of oath and protection.
\textbf{Thiasos:} Ledger, a warhorse with milk-white eyes.
\textbf{Hook:} Roran never takes ransom. He once spent three years in a debtor's cell. A stranger paid his price. He has been paying it forward ever since.

\begin{tabular}{ll}
\textbf{Body:} 4 & \textbf{Harm:} 6 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 4 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 2 (plate) \\
\textbf{Presence:} 3 & \textbf{Position:} Controlled \\
\end{tabular}

\textbf{Notable Rites:} \textit{Kindle Vow} (Low), \textit{Unbroken Vow} (Standard), \textit{Dawn's Judgment} (High)

\subsubsection{Hospitaller — Sister Bryn}

An old woman with a brass hook where her left hand should be. She wears a grey cloak and carries a small bronze bell. She tends a hospice at the edge of a village. She does not speak of her past.

\textbf{Codex:} The Grey Bell. Engraved with Rites of mercy and passage.
\textbf{Thiasos:} A hollowed bird with milk-white eyes, silent and watching.
\textbf{Hook:} Bryn was once a Daughter of the Cord. She cut a thread she should not have cut. The Cord has not forgotten.

\begin{tabular}{ll}
\textbf{Body:} 2 & \textbf{Harm:} 3 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 2 \\
\textbf{Spirit:} 4 & \textbf{Armor:} 0 \\
\textbf{Presence:} 3 & \textbf{Position:} Desperate (due to age) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Shared Bowl} (Low), \textit{Healing Waters} (Standard), \textit{Spiral's Witness} (Standard)

\subsubsection{Justicar — Elodie Venn}

A scarred woman in a dark violet cloak, her hand resting on a silver gavel at her belt. She travels with a spirit-hound that never blinks. She speaks in precedents.

\textbf{Codex:} The Silver Gavel. Engraved with Rites of law and judgment.
\textbf{Thiasos:} A spirit-hound named Precedent. Tracks oath-breakers.
\textbf{Hook:} Elodie once ruled against a baron who had taken a wetnurse's hand. The baron's lands were forfeit. The wetnurse's ghost still follows her.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 4 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 1 (cloak) \\
\textbf{Presence:} 4 & \textbf{Position:} Dominant (in court) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Consecrated Mandate} (Low), \textit{Seal the Covenant} (Standard), \textit{The Covenant Seal} (Standard)

\subsubsection{Witch-Hunter — Vigil-Commander Vorenus}

A gaunt man in an iron-mesh hood, carrying a brass censer that never goes cold. His lantern burns green. He hunts unlicensed magic and those who practice it.

\textbf{Codex:} The Censer of Unforgiving Flame. Engraved with Rites of purification.
\textbf{Thiasos:} A lantern that reveals hidden corruption. Its flame never wavers.
\textbf{Hook:} Vorenus once burned a village to catch a single heretic. He does not regret it. He dreams about it every night.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 1 (mesh hood) \\
\textbf{Presence:} 4 & \textbf{Position:} Desperate (when outnumbered) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Pure Flame} (Low), \textit{Cleansing Light} (Standard), \textit{Marked Prey} (Standard)

\subsubsection{Truth-Sworn — The Archivist}

A silent woman in grey robes, carrying a silver mirror in a leather case. She never speaks. She writes questions on a slate and shows the answers. Her memory-spirit records everything.

\textbf{Codex:} The Silver Mirror. Engraved with Rites of truth and memory.
\textbf{Thiasos:} A memory-spirit that records every word spoken in her presence.
\textbf{Hook:} The Archivist knows a secret about each member of the party. She has never told anyone. She will not say why.

\begin{tabular}{ll}
\textbf{Body:} 2 & \textbf{Harm:} 3 \\
\textbf{Wits:} 4 & \textbf{Fatigue:} 2 \\
\textbf{Spirit:} 4 & \textbf{Armor:} 0 \\
\textbf{Presence:} 2 & \textbf{Position:} Controlled (when writing) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Echoing Truth} (Standard), \textit{Immutable Record} (Standard), \textit{Unveiled Heart} (High)

\subsection{The Threshold}

\subsubsection{Warden — Sentinel of the Silent Stair}

A figure in grey armor standing motionless before a chapel door. Their face is hidden behind a helm. They do not eat, sleep, or speak. They wait.

\textbf{Codex:} The Shield of Silence. Engraved with Rites of warding.
\textbf{Thiasos:} A standing stone that wakes when they speak.
\textbf{Hook:} The Sentinel has guarded this door for forty years. No one knows what lies behind it. No one has asked.

\begin{tabular}{ll}
\textbf{Body:} 4 & \textbf{Harm:} 6 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 4 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 3 (plate + shield) \\
\textbf{Presence:} 2 & \textbf{Position:} Dominant (at the threshold) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Seal the Threshold} (Low), \textit{Warded Circle} (Standard), \textit{Consecrated Barrier} (High)

\subsubsection{Itinerant Monk — Brother Ash}

A wanderer in tattered robes, carrying a waystone compass that never points north. His road-dog walks beside him, one blue eye and one brown. He has no destination.

\textbf{Codex:} The Waystone Compass. Engraved with Rites of passage.
\textbf{Thiasos:} A road-dog named Path. Cannot be lost.
\textbf{Hook:} Brother Ash has walked the same road for thirty years. He does not remember why he started. He does not want to stop.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 4 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 0 \\
\textbf{Presence:} 2 & \textbf{Position:} Controlled (on the road) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Road-Sense} (Low), \textit{Waymark} (Standard), \textit{Traveler's Favor} (Standard)

\subsubsection{Death Monk — The Pallid Sister}

A grey-robed woman with a shepherd's crook. She appears at deathbeds, in battlefields, at the scene of accidents. She does not heal. She guides.

\textbf{Codex:} The Shepherd's Crook. Engraved with Rites of passage.
\textbf{Thiasos:} A spirit-shepherd that sings the names of the dead.
\textbf{Hook:} The Pallid Sister was once a midwife. She watched a mother die and a child live. She chose the child. The mother's ghost walks beside her.

\begin{tabular}{ll}
\textbf{Body:} 2 & \textbf{Harm:} 3 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 2 \\
\textbf{Spirit:} 4 & \textbf{Armor:} 0 \\
\textbf{Presence:} 3 & \textbf{Position:} Desperate (when guiding the dying) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Gentle Guide} (Low), \textit{Passing Flock} (Standard), \textit{Unburdened Farewell} (Standard)

\subsection{The Wild}

\subsubsection{Root-Knight — Elder Thorn}

A woman whose skin is the color of bark, whose hair grows leaves in spring. She carries a living staff that flowers when she is at peace.

\textbf{Codex:} The Living Staff. Carved with Rites of growth and season.
\textbf{Thiasos:} A wolf-ancestor that speaks in growls.
\textbf{Hook:} Elder Thorn was once a knight of a fallen kingdom. She swore an oath to defend a forest. The forest accepted. The kingdom is gone. She remains.

\begin{tabular}{ll}
\textbf{Body:} 4 & \textbf{Harm:} 6 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 4 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 1 (bark skin) \\
\textbf{Presence:} 2 & \textbf{Position:} Dominant (in forest) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Blood Offering} (Low), \textit{Fertile Death} (Standard), \textit{Cycle's Crown} (High)

\subsubsection{Grave-Warden — The Compost-Knight}

A figure in moss-covered armor, their face hidden behind a helm of woven roots. They carry a bone-chime that clicks in the wind.

\textbf{Codex:} The Bone-Chime. Engraved with Rites of decay and renewal.
\textbf{Thiasos:} A fungal familiar that speaks in spores.
\textbf{Hook:} The Compost-Knight was once a knight who failed to protect their lord. They sought the Carrion-King's mercy. The King granted it—by burying them in a compost heap. They rose three days later.

\begin{tabular}{ll}
\textbf{Body:} 4 & \textbf{Harm:} 6 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 4 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 2 (moss armor) \\
\textbf{Presence:} 1 & \textbf{Position:} Desperate (when separated from death) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Rotting Harvest} (Low), \textit{Borrowed Form} (Standard), \textit{Great Consumption} (High)

\subsubsection{Beast Warrior — Mkara of the Spiral Throw}

A Sekogo hunter with a leopard pelt over her shoulders. She carries three throwing knives and a predator's tooth on a cord. She does not smile.

\textbf{Codex:} The Predator's Tooth. Engraved with Rites of the hunt.
\textbf{Thiasos:} The spirit of the hunt. It does not speak. It watches.
\textbf{Hook:} Mkara killed her first leopard when she was twelve. She wore its skin. The leopard's spirit has been with her ever since.

\begin{tabular}{ll}
\textbf{Body:} 4 & \textbf{Harm:} 5 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 4 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 1 (leather) \\
\textbf{Presence:} 2 & \textbf{Position:} Dominant (when hunting) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Predator's Gaze} (Low), \textit{Shifting Shape} (Standard), \textit{Blood Scent} (Low)

\subsection{The Deep}

\subsubsection{Eldritch Knight — The Drowned Prince}

A figure in black glass armor that drips seawater. Their face is hidden behind a helm with no eye slits. They carry a shard of abyssal glass.

\textbf{Codex:} The Abyssal Glass. Engraved with Rites of pressure and silence.
\textbf{Thiasos:} A drowned spirit that speaks in bubbles.
\textbf{Hook:} The Drowned Prince was once a pirate king. He found a sunken temple and made a bargain with Khemesh. He does not remember his name. He does not need it.

\begin{tabular}{ll}
\textbf{Body:} 4 & \textbf{Harm:} 6 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 4 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 2 (glass) \\
\textbf{Presence:} 3 & \textbf{Position:} Desperate (in sunlight) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Crushing Whisper} (Low), \textit{Abyssal Grasp} (Standard), \textit{Sundering Deep} (High)

\subsubsection{Storm-Sworn — The Thunder Shaman}

A Ykrul warrior with lightning-scarred skin. She carries a drum made from storm-felled wood. Her hair stands on end when she is angry.

\textbf{Codex:} The Thunder Drum. Engraved with Rites of storm and shock.
\textbf{Thiasos:} A storm-spirit that crackles with lightning.
\textbf{Hook:} The Thunder Shaman was struck by lightning as a child. She survived. The storm has been following her ever since.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 4 & \textbf{Armor:} 0 \\
\textbf{Presence:} 3 & \textbf{Position:} Dominant (in storms) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Thunderous Presence} (Low), \textit{Fear-Eater} (Standard), \textit{World's Drum} (High)

\subsubsection{Tide Warrior — Captain Mira the Lucky}

A Sidhi pirate with a brass hook and a ship's wheel tattooed on her palm. She has one eye, one leg, and an uncanny ability to survive shipwrecks.

\textbf{Codex:} The Ship's Wheel. Engraved with Rites of tide and passage.
\textbf{Thiasos:} A selkie that sings in storms.
\textbf{Hook:} Mira has sunk seven ships. She has survived all seven. The sea owes her a debt. She is waiting to collect.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 1 (leather) \\
\textbf{Presence:} 4 & \textbf{Position:} Controlled (on a ship) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Wave-Tide} (Low), \textit{Changing Tide} (Standard), \textit{Storm-Queen's Hand} (High)

\subsection{The Shadow}

\subsubsection{Shadow Warrior — The Unseen}

A figure in a black feather cloak. Their face is hidden. Their voice is a whisper. They leave no footprints.

\textbf{Codex:} The Raven's Quill. Engraved with Rites of shadow and silence.
\textbf{Thiasos:} A raven-spirit that carries whispers.
\textbf{Hook:} The Unseen was once a spy for a fallen kingdom. They know where the bodies are buried. They know whose hands dug the graves.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 4 \\
\textbf{Wits:} 4 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 0 \\
\textbf{Presence:} 2 & \textbf{Position:} Dominant (in darkness) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Unlit Candle} (Low), \textit{Umbral Reservoir} (Standard), \textit{Umbral Dream} (High)

\subsubsection{Vowed Monk — The Silent Blade}

A warrior in grey robes, their mouth sewn shut with silver thread. They communicate with hand signs and written notes. Their blade never makes a sound.

\textbf{Codex:} The Sealed Lips Tattoo. Engraved with Rites of silence and witness.
\textbf{Thiasos:} A memory-spirit that speaks for them.
\textbf{Hook:} The Silent Blade swore an oath of silence after killing someone they loved. They have not spoken in twenty years. They do not plan to start.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 0 \\
\textbf{Presence:} 1 & \textbf{Position:} Controlled (when silent) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Gentle Seal} (Low), \textit{Weight of the Unsaid} (Standard), \textit{Cord of Witness} (Standard)

\subsubsection{Ronin — The Storm-Rider}

A masterless warrior on a grey horse. Their banner is torn. Their sword is chipped. They owe allegiance to no one.

\textbf{Codex:} The Storm-Banner. Engraved with Rites of freedom and storm.
\textbf{Thiasos:} A wind-spirit that rides beside them.
\textbf{Hook:} The Storm-Rider once served a lord who broke an oath. They cut their banner and rode away. They have been riding ever since.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 1 (leather) \\
\textbf{Presence:} 3 & \textbf{Position:} Dominant (on horseback) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Unlatched Step} (Low), \textit{Forgotten Road} (Standard), \textit{Storm's Refuge} (High)

\subsection{The Web}

\subsubsection{Web Warrior — The Spider-Blade}

A Lethai operative in a silk mask. Their movements are precise, silent, and impossibly fast. They carry a spider-silk cord that they can tie and untie with a thought.

\textbf{Codex:} The Spider-Silk Cord. Engraved with Rites of binding and mercy.
\textbf{Thiasos:} A spider-spirit that weaves invisible threads.
\textbf{Hook:} The Spider-Blade has killed nine targets. The tenth is their own reflection. They have been hunting it for years.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 4 \\
\textbf{Wits:} 4 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 0 \\
\textbf{Presence:} 3 & \textbf{Position:} Dominant (when unseen) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Hearth-Thread Knot} (Low), \textit{Strand of Inevitability} (Standard), \textit{Binding Knot} (High)

\subsubsection{Serpent Warrior — The Shed-Knight}

A warrior whose face changes each time they remove their helm. Their voice shifts. Their scars move. They are never the same person twice.

\textbf{Codex:} The Shed Fang. Engraved with Rites of transformation.
\textbf{Thiasos:} A serpent-spirit that speaks in sibilance.
\textbf{Hook:} The Shed-Knight has forgotten their original face. They are not sure they ever had one.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 1 (scale) \\
\textbf{Presence:} 3 & \textbf{Position:} Controlled (when masked) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Loosening Skin} (Low), \textit{Shedded Self} (Standard), \textit{Venom's Paradox} (Standard)

\subsubsection{Cursed Warrior — The Thorn-Sworn}

A figure covered in weeping scars. Their left hand is a brass hook. Their right hand holds a silver coin that never spends. They smell of brine and old grief.

\textbf{Codex:} The Silver Coin. Engraved with Rites of curse and bargain.
\textbf{Thiasos:} A Collector—a child with silver-wire eyes who watches in silence.
\textbf{Hook:} The Thorn-Sworn once served a baron who took their hand. They have been paying that debt ever since. They are almost done.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 4 & \textbf{Armor:} 0 \\
\textbf{Presence:} 2 & \textbf{Position:} Desperate (when the Collector watches) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Honeyed Curse} (Low), \textit{Binding Curse} (Low), \textit{False Dawn} (Standard)

\subsection{The Forge}

\subsubsection{Artificer — The Gear-Warrior}

A dwarf in a toolbelt, their left arm replaced with a clockwork prosthetic. They carry a schematic-slate that updates itself. They never stop tinkering.

\textbf{Codex:} The Schematic-Slate. Engraved with Rites of iteration and improvement.
\textbf{Thiasos:} A clockwork construct that chitters and ticks.
\textbf{Hook:} The Gear-Warrior built their own hand after losing it in a mining accident. They have been improving it ever since. It now has seventeen attachments.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 4 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 1 (clockwork arm) \\
\textbf{Presence:} 2 & \textbf{Position:} Controlled (when repairing) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Iterative Spark} (Low), \textit{Adaptive Mechanism} (Standard), \textit{Ultimate Iteration} (High)

\subsection{The Grave}

\subsubsection{Cairn-Knight — The Grave-Warden}

A warrior in rusted armor, standing motionless before a burial mound. Their axe is carved with runes. Their eyes are closed. They are not sleeping.

\textbf{Codex:} The Rune-Carved Axe. Engraved with Rites of wergild and memory.
\textbf{Thiasos:} An ancestral shade that fights beside them.
\textbf{Hook:} The Grave-Warden is guarding their own grave. They died forty years ago. They did not stay dead.

\begin{tabular}{ll}
\textbf{Body:} 4 & \textbf{Harm:} 6 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 4 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 2 (rusted plate) \\
\textbf{Presence:} 2 & \textbf{Position:} Dominant (at the cairn) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Crossing Fee} (Low), \textit{Ancestral Consultation} (Low), \textit{Mediated Pact} (Standard)

\subsubsection{Bloodsworn — The Cord-Knight}

A Daughter of the Cord with Rites tattooed on every inch of her skin. Her eyes are milk-white. A hollowed bird perches on her shoulder.

\textbf{Codex:} Her skin. Engraved with Rites of debt and witness.
\textbf{Thiasos:} A hollowed bird with milk-white eyes.
\textbf{Hook:} The Cord-Knight has been paying her grandmother's debt for forty years. She does not know what the debt was. She only knows it is almost paid.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 2 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 4 & \textbf{Armor:} 0 \\
\textbf{Presence:} 3 & \textbf{Position:} Desperate (when the Cord calls) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Milk Teeth} (Low), \textit{Unseen Midwife} (Standard), \textit{Unnamed} (High)

\subsection{The Fracture}

\subsubsection{Fractal Warrior — The Paradox-Sworn}

A warrior whose eyes see angles that should not exist. Their Codex is a broken mirror. They move in ways that make observers dizzy.

\textbf{Codex:} The Broken Mirror. Engraved with Rites of paradox and pattern.
\textbf{Thiasos:} A pattern-spirit that speaks in equations.
\textbf{Hook:} The Paradox-Sworn saw their own death in a fractured mirror. They have been trying to avoid it ever since. The mirror shows a different death each time.

\begin{tabular}{ll}
\textbf{Body:} 2 & \textbf{Harm:} 4 \\
\textbf{Wits:} 4 & \textbf{Fatigue:} 2 \\
\textbf{Spirit:} 4 & \textbf{Armor:} 0 \\
\textbf{Presence:} 2 & \textbf{Position:} Controlled (when predicting) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Ninth Figure} (Low), \textit{Hypercognitive Lattice} (Standard), \textit{Ninth Revelation} (High)

\subsubsection{Moon Knight — The Silver-Sworn}

A warrior in pale armor who fights only under moonlight. Their Codex is a silver disc. They carry no weapon—the moon fights for them.

\textbf{Codex:} The Silver Disc. Engraved with Rites of moon and shadow.
\textbf{Thiasos:} A shadow-cat that walks silently beside them.
\textbf{Hook:} The Silver-Sworn was cursed to become a wolf under the full moon. They made a bargain with Lunara. Now they hunt werewolves instead.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 5 \\
\textbf{Wits:} 3 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 3 & \textbf{Armor:} 1 (moonlight) \\
\textbf{Presence:} 2 & \textbf{Position:} Dominant (under moonlight) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Moonlit Mirror} (Low), \textit{Triple Road} (Standard), \textit{Lunar Conjunction} (Standard)

\subsection{The Lethai Ghe'hai}

\subsubsection{Spec Ops — The Silk-Blade}

An elite operative with no name, no face, and no past. They wear a woven silk mask. Their movements are precise, efficient, and forgettable.

\textbf{Codex:} The Woven Mask. Engraved with Rites of silk and silence.
\textbf{Thiasos:} A spider-spirit that weaves invisible threads.
\textbf{Hook:} The Silk-Blade has completed seventeen missions for the Lethai courts. They do not remember any of them. The mask remembers for them.

\begin{tabular}{ll}
\textbf{Body:} 3 & \textbf{Harm:} 4 \\
\textbf{Wits:} 4 & \textbf{Fatigue:} 3 \\
\textbf{Spirit:} 2 & \textbf{Armor:} 0 \\
\textbf{Presence:} 3 & \textbf{Position:} Dominant (when masked) \\
\end{tabular}

\textbf{Notable Rites:} \textit{Silk Binding} (Low), \textit{Mask of Many Faces} (Standard), \textit{Unspoken Question} (High)

The thread held. The water carried. That is not a Rite. That is grace.
