\chapter{Managing Resources}
\label{ch:gm-resources}
\index{resources!GM guidance}

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title={Philosophy of Resources}]
\index{resources!philosophy}
\textbf{Resources are not there to hoard – they are tools to create drama.}

\begin{itemize}
  \item \textbf{Spend Boons freely.} Hoarding them for XP conversion starves the table of tension and robs players of their safety net.
  \item \textbf{Let Fatigue build.} Fatigue is not punishment; it is the rising tide that forces hard choices and heroic comebacks.
  \item \textbf{Obligation is plot, not punishment.} When a Patron calls in a debt, a story hook arrives. Embrace it.
\end{itemize}

The most memorable sessions come from characters spending their last Boon on a desperate re-roll, staggering forward under three Fatigue, or accepting a Patron’s demand that changes everything.
\end{tcolorbox}

In \textbf{Fate's Edge}, resources are narrative levers, not inventory lists. From the last sip of water in sun‑scorched wastes to the fragile loyalty of a Ykrul war‑band, every resource tells a story, and every story demands its price. As GM, you are the keeper of these threads—the weaver of scarcity and abundance. This chapter shows how to manage the systems that fuel mortal ambitions and epic campaigns.

\section{Supply Clock: The Pulse of Survival}
\label{sec:supply-clock}
\index{Supply Clock}

The \textbf{Supply Clock} tracks the party’s access to essentials: food, water, gear, and logistical support. It is a shared party resource—a narrative lever that tightens tension when characters are isolated, pursued, or cut off from civilization.

\subsection{Supply Clock States}
\label{subsec:supply-states}

\begin{tcolorbox}[title=Supply Clock Conditions, colback=white!97!gray, colframe=black!80!gray]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Segments Filled} & \textbf{Narrative Effects} \\
\midrule
0 (Full) & Well‑equipped and prepared; no penalty. \\
1 (Low) & Minor complications: bland rations, damaged arrows, thinning waterskins. \\
2 (Dangerous) & Each character gains \textbf{Fatigue 1} as scarcity takes its toll. \\
3 (Empty) & Severe penalties: starvation, dehydration, equipment failure imminent. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\subsection{When to Advance the Supply Clock}
\label{subsec:supply-advance}

The Supply Clock advances when the world turns against the party’s preparations:
\begin{itemize}
  \item Extended travel through hostile lands without proper provisioning (see \textbf{Travel Framework}, Chapter 9 of the Player’s Guide or SRD Section 41).
  \item The GM spends \textbf{2+ Story Beats} on logistics failures or environmental hardships.
  \item The party chooses to travel light for speed or stealth.
  \item Failed \textbf{Survival} or \textbf{Craft} rolls related to hunting, foraging, or repair.
\end{itemize}

\subsection{Clearing Supply Segments}
\label{subsec:supply-clear}

The clock recedes when the party finds respite:
\begin{itemize}
  \item Reaching civilization resets the clock to \textit{Full}.
  \item A group \textbf{Survival} check (Wits + Survival, DV 2) in favorable conditions clears 1 segment.
  \item Downtime spent in relative safety removes 1 segment.
  \item Successful provisioning (a good hunt, a discovered cache) can reduce segments.
\end{itemize}

\paragraph{Example.}
A week‑long sea passage with uncertain winds. A failed \textbf{Navigation} roll causes the GM to spend 2 SB – filling two segments as supplies spoil. The party is now at \textit{Low Supply}. A second failed roll fills another segment – \textit{Dangerously Low}. Fatigue sets in. The sea gnaws at their endurance.

\section{Fatigue: The Weight of the World}
\label{sec:fatigue-gm}
\index{Fatigue!GM guidance}

Fatigue represents the cumulative toll of journeying through a world that rarely offers comfort: exhaustion that seeps into bones, hunger that hollows cheeks, strain that clouds judgment.

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title={The Roller‑Coaster Effect}]
\index{Fatigue!roller-coaster effect}
\textbf{Taking Harm clears Fatigue temporarily} – the shock of a wound can jolt a character back into action. This creates a natural “roller coaster” of tension:

\medskip
\noindent\textit{Example:} Kael is at Fatigue 3 (Desperate, –2 dice). A goblin stabs him for Harm 1. The Harm conversion procedure: first convert Harm → Fatigue (none, light armor), then apply remaining Harm – which clears all Fatigue. Kael’s Fatigue drops to 0, but his Harm increases to 1. He is wounded but suddenly clear‑headed, ready to fight back.

\medskip
Use this effect to create dramatic swings: a character near collapse who takes a wound and finds a second wind. It keeps combat unpredictable and rewards players who embrace risk.
\end{tcolorbox}

\subsection{Fatigue Effects}
\label{subsec:fatigue-effects}

Each point of Fatigue worsens the character’s \textbf{Position} by one step (Dominant → Controlled → Desperate). If already Desperate, each additional Fatigue instead imposes a \textbf{–1 die penalty} on all rolls. Penalties from Fatigue and Harm \textbf{stack}.

\begin{tcolorbox}[title=Fatigue Progression, colback=white!97!gray, colframe=black!80!gray]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Fatigue Level} & \textbf{Effect} \\
\midrule
1 & Position worsens one step (Dominant → Controlled → Desperate) \\
2 & If already Desperate, –1 die; otherwise Position worsens \\
3 & –2 dice (if already Desperate from Fatigue) \\
4+ & Collapse, KO, or spiritual break (incapacitated) \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\subsection{Clearing Fatigue}
\label{subsec:fatigue-clear}

Recovery requires genuine respite:
\begin{itemize}
  \item A \textbf{short rest} (quiet watch, food/water) removes \textbf{2 Fatigue}.
  \item A \textbf{full night’s rest} in safety removes \textbf{all Fatigue}.
  \item Fatigue cannot be removed while the party is at \textit{Dangerous} or \textit{Empty} Supply.
  \item Medical attention (\textbf{Presence + Medicine}, DV 2) during downtime can remove an additional 1 Fatigue.
\end{itemize}

\paragraph{Narrative note:}
Fatigue is not only physical – it can reflect mental strain, grief, or spiritual exhaustion. A failed ritual might leave a caster at Fatigue 2 from metaphysical backlash alone.

\subsection{Fatigue → Harm Conversion}
\label{subsec:fatigue-harm}

Each character has a Fatigue track equal to their \textbf{Body} attribute. When the track fills:
\begin{enumerate}
  \item Increase \textbf{Harm} by one level (0→Harm 1, Harm 1→Harm 2, Harm 2→Harm 3).
  \item Clear all Fatigue.
\end{enumerate}
This conversion can occur multiple times in a scene. (See the \textit{Player’s Guide} for full Harm rules.)

\section{Boons: The Currency of Resilience}
\label{sec:boons-gm}
\index{Boons!GM guidance}

Boons are narrative tokens earned by embracing risk. They reward \textit{failure with opportunity} – the silver lining in clouds of defeat.

\begin{tcolorbox}[colback=yellow!5!white, colframe=yellow!75!black, title={Why You Should Spend Boons – The Math of Fun}]
\index{Boons!spending encouragement}
\textbf{Hoarding Boons for XP conversion leads to less engagement and lower fun scores.}

\medskip
\noindent Every Boon you hold is a dramatic moment waiting to happen:
\begin{itemize}
  \item Re‑rolling a critical failure turns despair into triumph.
  \item Improving Position before a risky roll turns a Desperate gamble into a Controlled chance.
  \item Activating an Asset brings your character’s network into the story.
\end{itemize}
\textbf{XP from Boons is a trickle (max 2 per session).} The real reward is \textit{agency}. The player who spends Boons shapes the story; the player who hoards them watches from the sidelines.

\medskip
\noindent As GM, remind players: “You have 3 Boons. What will you spend them on?” When a player sits on 5 Boons, gently suggest they use one before the scene ends. A spent Boon is a moment of player power. An unspent Boon is wasted potential.
\end{tcolorbox}

\subsection{Earning Boons (GM Reminders)}
\label{subsec:boons-earn}

\begin{itemize}
  \item On a \textbf{Partial} (1 Boon) or \textbf{Miss} (2 Boons) with meaningful consequences.
  \item Through clever or risky roleplay that drives the story.
  \item Via bond‑driven actions with intricate descriptions.
  \item At GM discretion for exceptional collaborative play.
\end{itemize}
\textbf{Limit:} At most \textbf{2 Boons} from failures per character per scene. Repetition of the same approach with the same stakes does not award additional Boons.

\subsection{Boon Economy (Quick Reference)}
\label{subsec:boons-economy}

\begin{itemize}
  \item \textbf{Holding cap:} Maximum 5 Boons.
  \item \textbf{Carryover:} At the end of each scene, reduce held Boons to a maximum of 2 (excess lost).
  \item \textbf{Conversion:} Once per session during downtime, convert 2 Boons → 1 XP (max 2 XP per session via conversion).
\end{itemize}

\subsection{Spending Boons (Player Options)}
\label{subsec:boons-spend}

Players may spend 1 Boon to:
\begin{itemize}
  \item Re‑roll a single die (after seeing the pool).
  \item Activate an on‑screen Asset.
  \item Improve Position by one step before a roll.
  \item Power certain Rites or abilities as described in their text.
\end{itemize}

\paragraph{Substituting Fatigue for Boons (Optional)}
If a player has no Boons remaining, they may substitute \textbf{1 Fatigue} for 1 Boon \textit{once per scene}. This represents pushing through exhaustion for a last chance. The GM may veto if the fiction does not support it.

\subsection{Anti‑Fishing Measures (GM Guidance)}
\label{subsec:boons-anti-fishing}

\begin{itemize}
  \item \textbf{Once/Scene Cap:} At most 2 Boons from failures per character per scene.
  \item \textbf{Repetition Rule:} Same approach + same stakes in the same scene cannot award another Boon.
  \item \textbf{Position Gate:} \textit{Dominant} tests with trivial fallout do not award Boons.
\end{itemize}
\emph{Design note:} Boons reward narrative risk, not repeated failure.

\section{Followers and Assets: Power Beyond the Self}
\label{sec:followers-assets-gm}
\index{Followers!GM guidance}\index{Assets!GM guidance}

Players can invest XP in \textbf{Followers} (on‑screen allies) and \textbf{Assets} (off‑screen resources). These are story agents with their own motivations and risks.

\subsection{Followers: On‑Screen Allies}
\label{subsec:followers-gm}

Followers are bought with XP and tracked by a \textbf{Cap} (maximum assist bonus). Cost = Cap\(^2\) XP; downtime 1–3 days to recruit.

\paragraph{Assisting in Scenes}
\begin{itemize}
  \item Assist dice equal \(\min(\text{Cap}, \text{helper’s relevant Skill})\).
  \item Total assist from all sources capped at \textbf{+3 dice} (the \textit{Exceptional Coordination} talent allows +4 from one follower).
  \item Only one follower may assist a given action.
\end{itemize}

\paragraph{Follower Independent Actions}
Once per scene, \textbf{one} follower may take a small independent action (the party can use this ability only once per scene, not per follower). Examples:
\begin{itemize}
  \item \textbf{Scout \& Signal} – Change an ally’s next action to \textit{Dominant}.
  \item \textbf{Distract \& Draw} – Reduce a threat/hunt/escape clock by 1 tick.
  \item \textbf{Fetch \& Carry} – Move a small object through danger.
\end{itemize}
\textbf{Cost:} Mark \textbf{Exposure +1} or \textbf{Harm 1} on that follower.

\paragraph{Followers as Successors}
A Follower is not just a helper – they can become a \textbf{successor} if the player’s main character retires, dies, or ascends. This is the core of the Legacy Engine (see \ref{ch:legacy-engine} in the full GM Guide). When a player’s character reaches a narrative endpoint, a trusted Follower (Cap 3+) can step into the spotlight with inherited Bonds, Assets, and a reduced XP cost.

\paragraph{When to Invest in Followers (GM Guidance)}
\begin{tcolorbox}[title=When to Invest in Followers, colback=white!97!gray, colframe=black!80!gray]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Player Archetype} & \textbf{Investment Advice} \\
\midrule
Solo & Minimal followers (Cap 1–2). Use for specific utility (scout, healer). \\
Mixed & One Cap 2–3 follower. The sweet spot for most players. \\
Mastermind & Cap 3–5 followers. Build a network. Accept that they will be endangered. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\paragraph{Follower Upkeep and Risk}
\begin{itemize}
  \item Each downtime, pay XP equal to Cap \textbf{or} spend a scene tending the relationship.
  \item Neglect → follower becomes \textit{Wary} (or \textit{Seized} if already Wary).
  \item If the GM spends \textbf{2+ SB} on an action you take with assistance, they may mark Exposure or Harm on the follower instead of other consequences (fiction first).
  \item \textbf{Off‑Screen Capability:} Once per downtime, a Cap 3+ follower can solve a significant problem but generates \textbf{1 SB} for the party (the GM describes the resulting complication).
\end{itemize}

\paragraph{Follower Conditions}
Followers track two meters:
\begin{description}
  \item[Exposure] – Attention, stress, narrative pressure.
  \item[Harm] – Injury, trauma, damage.
\end{description}
State: \textit{Maintained} (full) → \textit{Neglected} (–1 die) → \textit{Compromised} (unavailable).

\subsection{Assets: Off‑Screen Influence}
\label{subsec:assets-gm}

Assets are titles, safehouses, spy rings, charters. They do not act in scenes directly but change the fiction between sessions.

\begin{tcolorbox}[title=Asset Tiers, colback=white!97!gray, colframe=black!80!gray]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Tier} & \textbf{Cost and Examples} \\
\midrule
Minor & 4 XP (1 day) – safehouse, small shop, local contact. \\
Standard & 8 XP (1 week) – noble title, spy ring, workshop. \\
Major & 12 XP (1 month) – city license, fortress lease, regional network. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\paragraph{Using Assets}
\begin{itemize}
  \item \textbf{Off‑Screen Effect:} Use once per session for free (the quiet work between adventures).
  \item \textbf{On‑Screen Activation:} Spend 1 Boon to use the asset dramatically in a scene.
  \item \textbf{Downtime Activation:} Pay 2 XP or 1 Boon to activate an asset at campaign start or during downtime.
  \item The asset must have plausible scope for the intended effect.
\end{itemize}

\paragraph{Asset Upkeep}
Each downtime:
\begin{itemize}
  \item \textbf{Efficient (higher XP, less time):} Cost = \(\max(1, \frac{\text{Acquisition XP}}{3})\). Minimal effort.
  \item \textbf{Intensive (lower XP, more time):} Cost = 1 XP. Requires a dedicated downtime action with significant personal involvement.
  \item Failure to pay → asset becomes \textit{Neglected} (or \textit{Compromised} if already Neglected).
\end{itemize}

\paragraph{Asset Condition}
\begin{description}
  \item[Maintained] – Full capability.
  \item[Neglected] – –1 die when used; requires attention.
  \item[Compromised] – Unavailable until repaired or recovered.
\end{description}

\section{Terrestrial Patrons: Favours and Debts}
\label{sec:terrestrial-patrons}
\index{Patrons!terrestrial}\index{Debt Clock}

Not every patron is a cosmic entity. Powerful mortals—nobles, crime lords, generals, merchant princes—can act as \textbf{terrestrial patrons}. They offer tangible aid (escorts, legal protection, supplies, even ongoing Assets like a trade route or a safehouse) in exchange for favours.

\subsection*{Assets Without XP Cost, With Obligation}
A terrestrial patron can provide the same benefits as a Minor or Standard Asset (see \ref{sec:assets-gm}), but the party does not pay XP for it. Instead, they accept a \textbf{Debt Clock [4]} (or larger, for major favours) and track Obligation to the patron just as they would to a supernatural Patron (see \ref{sec:obligation-capacity}). The patron’s “Rites” are favours; the “Obligation” is the growing weight of what the party owes.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title={Asset or Patron? A GM’s Shortcut}]
If the party needs a recurring advantage (a warehouse, a guild contact, a ship) and you want to avoid bookkeeping, offer it through a \textbf{terrestrial patron} instead of an XP‑cost Asset. The party gets the benefit immediately; the price is story fuel (Obligation). If the players prefer to own the Asset outright and handle its upkeep themselves, they can purchase it with XP as usual. Both paths are valid – choose the one that generates the most interesting complications for your table.
\end{tcolorbox}

\subsection{How Terrestrial Patrons Work}
When a character accepts a favour from a terrestrial patron, start a \textbf{Debt Clock} with the patron’s name (e.g., “Prefect Marcellus’s Favour [4]”). Mark one segment each time the character accepts a significant favour (e.g., a bribe, a military escort, access to a restricted archive). When the clock fills, the patron calls in the debt – a mission to be performed, a political favour, or a direct demand.

The party also tracks a separate \textbf{Obligation} tally for each terrestrial patron, exactly as they would for a supernatural Patron (see \ref{sec:obligation-capacity}). Each accepted favour typically adds +1 Obligation. When Obligation exceeds the character’s (or party’s) capacity, they mark Fatigue and the patron may send a reminder – a messenger, a withdrawn favour, or a public hint that the debt is due.

Refusing to honour the debt turns the patron into an enemy (or at least a reliable antagonist). The GM may advance associated \textit{Crisis} clocks or introduce complications from the patron’s faction.

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title={Use Debt Clocks Aggressively}]
\index{Debt Clock!GM advice}
\textbf{Debt clocks are story fuel.} Every time a player accepts a favour, they are inviting future conflict. Don’t let debts linger – call them in at the worst possible moment:

\begin{itemize}
  \item \textbf{Early (1–2 segments):} A messenger arrives with a “reminder.” The patron asks for a small favour (deliver a package, spread a rumour).
  \item \textbf{Mid (3–4 segments):} The patron demands a significant service (escort a caravan through dangerous territory, provide alibis for a crime).
  \item \textbf{Full (5–6 segments):} The patron calls in the debt at a critical moment – during a negotiation, a battle, or a ritual. Refusal means the patron becomes an enemy.
\end{itemize}

Debts that are never called are wasted potential. Be bold.
\end{tcolorbox}

\subsection{Examples of Terrestrial Patron Debts}
\begin{tcolorbox}[title=Sample Terrestrial Patrons, colback=white!97!gray, colframe=black!80!gray]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Patron} & \textbf{Typical Favours and Debt Trigger} \\
\midrule
Prefect Marcellus (Ecktorian commander) & Grants safe passage, food, and a squad of guards. Debt: demand a dangerous mission behind enemy lines. \\
Censor Cassia (judicial authority) & Offers legal immunity, archive access. Debt: require the party to expose a political rival. \\
Khatun Sarnai (Ykrul chieftain) & Provides remounts, safe camp, steppe guides. Debt: demand a blood‑price or a raid on a rival clan. \\
Tema (Vilikari merchant) & Smuggles goods, offers black market contacts. Debt: require the party to protect a secret caravan. \\
\addlinespace
\textbf{Additional Examples} & \\
Mistress Ilvar (Silkstrand crime lord) & Grants access to the Black Bazaar and a crew of cutpurses. Debt: demand a public assassination during the Exchange opening. \\
Abbot Valerius (Thepyrgos Synod) & Provides sanctuary for fugitives and letters of indulgence. Debt: demand the party retrieve a “heretical” text from a sealed archive – without being caught. \\
Salt-Prince Kor (Zakov) & Gives a letter of marque and a fast cutter. Debt: demand the party act as decoys to draw away a rival’s fleet. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\subsection{Running Terrestrial Patrons}
\begin{itemize}
  \item Track each patron’s Debt Clock and the party’s Obligation to them on a visible handout.
  \item When a player asks for a favour, tell them the cost: “Marcellus will give you the keys to the armory, but that’s worth one segment and +1 Obligation. Are you willing?”
  \item When a Debt Clock fills, the favour is due. Give the players a hook that fits the patron’s personality and resources.
  \item If the party refuses, tick the associated \textit{Crisis} clock or introduce a complication (e.g., loss of safe passage, arrest warrant, hired thugs). The patron may also demand immediate payment of the Obligation in the form of a significant sacrifice or a public apology.
\end{itemize}

\subsection{Optional: Terrestrial Patron Reputation}
Some campaigns may track a separate \textbf{Reputation Track} for each patron. High reputation grants better terms (more favours before the debt clock fills). Low reputation may cause the patron to demand immediate payment before offering any help. Reputation can be advanced by completing the patron’s tasks on time or by publicly supporting their cause.

\subsection{Clearing Terrestrial Obligation}
Obligation to a terrestrial patron is cleared the same way as Obligation to a supernatural Patron: by performing an Act of Service (see \ref{sec:acts-of-service}) during downtime. The service must be concrete – repaying a loan, performing a public favour, or delivering a rival’s secret. The GM sets the scope: a small favour clears 1–2 segments; a major undertaking may clear the entire clock.

\subsection{Cross‑Reference}
\begin{itemize}
  \item For player‑facing rules on terrestrial patrons, see \ref{sec:player-terrestrial-patrons} (Player’s Guide).
  \item For Asset upkeep and the XP alternative, see \ref{sec:assets-gm} and \ref{subsec:assets-choice}.
  \item For Obligation mechanics, see \ref{sec:obligation-capacity}.
\end{itemize}

% The labels below are kept for cross‑linking from other parts of the guide
\label{sec:terrestrial-patrons-gm}

\subsection{Milestones}
\label{subsec:xp-milestones}

At the end of a major story arc:
\begin{itemize}
  \item +8–12 XP to all players.
  \item +2 XP bonus to one player for a signature moment of the arc.
\end{itemize}

\section{Campaign Resources: Mandate and Crisis}
\label{sec:campaign-resources}
\index{Mandate}\index{Crisis}

Two campaign‑level clocks track the party’s influence and the world’s resistance.

\subsection{Mandate Clock (0–6)}
\label{subsec:mandate}

Tracks the party’s public legitimacy.
\begin{itemize}
  \item \textbf{High Mandate:} Allies seek them out, resources flow, doors open.
  \item \textbf{Low Mandate:} Suspicion, bureaucratic obstacles, support withers.
\end{itemize}
Advance the clock when the party gains renown (clean victories, public accolades) or loses face (scandals, broken oaths).

\subsection{Crisis Clock (0–6)}
\label{subsec:crisis}

Tracks the opposition’s growing strength.
\begin{itemize}
  \item \textbf{Rising Crisis:} Complications escalate, enemies grow bolder.
  \item \textbf{Managed Crisis:} Breathing room, opportunities to strike back.
\end{itemize}
Advance the clock when the party fails key objectives, ignores looming threats, or the GM spends SB on faction escalation.

\paragraph{Example.}
If the party brokers a peace treaty, advance \textit{Mandate} by 1. If they assassinate a rival leader but leave evidence, advance \textit{Crisis} by 1. When either clock reaches 6, trigger a major faction response (a summons, an assassination attempt, a public trial).

\section{Combat Resource Management}
\label{sec:combat-resources}
\index{combat resources!GM guidance}

In combat, resource management takes on desperate urgency.

\subsection{Supply in Combat}
\label{subsec:combat-supply}

\begin{itemize}
  \item \textbf{Intense Combat:} GM may spend 1 SB to fill 1 Supply segment (arrows break, waterskins puncture).
  \item \textbf{Prolonged Engagement:} Each hour of sustained combat adds 1 Supply segment.
  \item \textbf{Ammunition Depletion:} Ranged weapons may run low, requiring scavenging actions (e.g., \textbf{Craft} roll to recover arrows).
\end{itemize}

\subsection{Fatigue in Combat}
\label{subsec:combat-fatigue}

\begin{itemize}
  \item Characters apply Fatigue penalties as described in \ref{sec:fatigue-gm}.
  \item Reaching Fatigue level 4+ during combat causes immediate collapse (incapacitated).
  \item Fatigue cannot be cleared during active combat.
\end{itemize}

\subsection{Follower Combat Integration}
\label{subsec:combat-followers}

\begin{itemize}
  \item Followers assist using their Cap (see \ref{subsec:followers-gm}).
  \item Spending \textbf{2+ SB} in combat can endanger an assisting follower (mark Exposure or Harm).
  \item Followers can take combat‑relevant independent actions (cost: Exposure +1 or Harm 1).
  \item After a follower acts twice in a high‑risk combat scene, mark Exposure +1.
\end{itemize}

\subsection{Tactical Clocks}
\label{subsec:tactical-clocks}

Tactical clocks represent persistent combat conditions that drain party resources.

\begin{tcolorbox}[title=Tactical Clock Effects, colback=white!97!gray, colframe=black!80!gray]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Clock} & \textbf{Resource Drain} \\
\midrule
Mob Overwhelm [6] & Accelerates Supply depletion and Fatigue. \\
Fatigue Spiral [4] & Accelerates existing Fatigue, makes each action harder. \\
Morale Collapse [6] & Generates SB, reduces effectiveness. \\
Environmental Collapse [8] & Creates new Supply and safety concerns. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\subsection{Over‑Stack Trigger}
\label{subsec:overstack-trigger}
\index{combat!over-stack trigger}

\begin{itemize}
  \item If the party \textbf{enters} a scene with \(\geq 3\) \textbf{structural advantages}, the GM \textit{may} apply Over‑Stack once for that scene.
  \item Choose: start one obstacle at +1 DV, \textbf{or} bank +1 SB for the first Deck Twist.
  \item Over‑Stack applies only to advantages held at scene framing, not those generated through play.
\end{itemize}

\paragraph{Examples of Structural Advantages}
Active buffs, favourable terrain (high ground, chokepoint), unused Follower Initiative, on‑screen Asset activation, enemy disadvantage (surprised, trapped, outnumbered), ritual preparation that applies to the scene.

\section{Narrative First: The Fiction Is the Ledger}
\label{sec:resources-narrative-first}
\index{narrative first!resources}

In Fate’s Edge, arrows, rations, and waterskins are tracked only in the fiction that surrounds them. Mechanics engage only when those resources become scarce enough to matter. The focus remains always on \textbf{narrative tension} – the gnawing hunger, the fading light, the last arrow – not sterile bookkeeping.

\paragraph{Abstracting Coin and Time.}
Routine expenses and petty bribes are handled narratively (“You have enough coin from past jobs to cover a week’s stay”). For major expenses (bribing a corrupt official, outfitting an expedition, buying a rare magical component), players may spend \textbf{XP} in place of coin (1 XP = a substantial sum). Use the “XP as currency” rule sparingly; it should feel like a meaningful sacrifice, not a tax.

\paragraph{GM Reminder.}
When a player asks “Do I have enough arrows?” the answer is almost always “yes” until it becomes dramatically interesting to say “no.” Similarly, track Supply only when the party is isolated, pursued, or in a hostile environment. Let the world breathe with its own needs and abundances. And when the dice say the world pushes back against mortal plans – \textbf{listen to what they tell you about the price of ambition.}