\chapter{Campaigns, Clocks, and Consequences}
\label{ch:campaigns-clocks}
\index{campaigns}\index{clocks}\index{consequences}
\index{travel!GM guidance}

In \textbf{Fate's Edge}, campaigns are not a sequence of disconnected adventures—they are \textbf{pressure systems}.  
Every choice the players make pushes on something: a faction, a threat, a relationship, a belief. Over time, those pressures accumulate, shift, and eventually \emph{break}.

Your job as GM is not to hide this pressure—it is to \textbf{shape it, reveal it, and let the players decide where it lands}.

This chapter presents the core tools for doing that work: \textbf{Clocks}, \textbf{Consequences}, and long‑term campaign structures. Used well, they create momentum without railroading and stakes without arbitrary escalation.

% =========================================================
% NEW: TRAVEL DESIGN FOR GMS
% =========================================================
\section{Running Travel as a Medieval Trek}
\label{sec:gm-travel}
\index{travel!GM techniques}

Travel in Fate’s Edge is not a line on a map. It is a **medieval or Renaissance journey** – uncertain, weather‑beaten, dependent on local guides, hospitality, and the kindness (or cruelty) of strangers. Use the procedures from the \textit{Travel Guide} and the \textit{Player’s Guide} (Chapter 10) as your skeleton, then layer on the following GM techniques to make every leg feel like a story.

\subsection{Lead with Weather, Not Distance}
Open each leg with a weather description, not a coordinate.  
\textit{“The morning mist is so thick you can’t see the second wagon. The guide says it’ll burn off by noon – if the wind shifts.”}  
Then set Position based on that weather (e.g., Desperate for navigation, Controlled for stealth). The weather \emph{is} the terrain.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Weather as Position Modifier]
  \begin{itemize}
    \item \textbf{Favorable} (tailwind, clear skies): +1 Position to sailing/stealth maneuvers.
    \item \textbf{Adverse} (headwind, fog): –1 Position to all sailing actions.
    \item \textbf{Severe} (storm, whiteout): –1 Position and +1 DV to all actions.
  \end{itemize}
  For a full weather table, see the \textit{Travel Guide} or roll d12 on the \textbf{Weather Engine} (GM Appendix).
\end{tcolorbox}

\subsection{Maps Are Lies; Guides Are Truth}
In a medieval trek, you don't have GPS. You have a local who knows which ford is passable, which village sells stale bread, and which innkeeper is a cutpurse.  
\textbf{Mechanic:} Require a \textbf{Guide} role for each leg (a PC or hired NPC). The Guide’s roll (Wits + Survival or Lore) determines if they find the safe route, the short route, or the one that leads to a cultist ambush. If the Guide fails, the leg adds a \textbf{Complication} (e.g., “The bridge is out – now you have to ford the river at night”).

\subsection{Hospitality as a Resource}
Inns, monasteries, and farmsteads are not just “rest stops.” They are \textbf{social encounters} with their own risks and rewards.  
\textbf{Mechanic:} When the party seeks shelter, draw a \textbf{Heart} from the regional deck. That’s the host. Their attitude (Warm, Suspicious, Greedy) affects whether you get a hot meal, a dry bed, or a knife in the back. Spending a \textbf{String} (e.g., “I helped this village last year”) can improve the outcome.

\subsection{Road Conditions Change Daily}
A road that was passable yesterday may be a mud pit today.  
\textbf{Mechanic:} At the start of each leg, choose or roll a \textbf{Road Condition} from the table below. Apply its effect for the entire leg.

\begin{fatebox}[Road Conditions (d6)]
  \begin{tabularx}{\textwidth}{lX}
    \toprule
    \textbf{d6} & \textbf{Condition \& Effect} \\
    \midrule
    1 & \textbf{Dry \& Hard} – Travel Clock advances +1 segment on any success. \\
    2 & \textbf{Muddy} – Supply ticks –1 on any failure (wagons stick, horses struggle). \\
    3 & \textbf{Washed Out} – Detour adds an extra scene (side quest or delay). \\
    4 & \textbf{Icy} – Desperate Position for all movement rolls. \\
    5 & \textbf{Swollen Rivers} – Fords become gambling; any failure ticks Supply –1 and adds a Delay. \\
    6 & \textbf{Dust Storm / Fog} – Navigation DV +1; ranged attacks –1 die. \\
    \bottomrule
  \end{tabularx}
\end{fatebox}

\subsection{Regional Decks as Story Engines}
The \textit{Travel Guide} provides regional decks (Spade=place, Heart=actor, Club=pressure, Diamond=leverage). Draw one of each at the start of a leg and let them suggest the next scene. Do not prep beyond the cards.  
\textbf{Example:} For a leg in the Mistlands, you draw:  
- Spade: \textit{Bell‑Line levee} – a raised causeway with iron bells.  
- Heart: \textit{Mist‑Seer} – a blind oracle who speaks in riddles.  
- Club: \textit{Witchlight Count} – will‑o’‑wisps that force a memory test.  
- Diamond: \textit{Ferry Token} – a pass that guarantees safe crossing.  

You now have a scene: the party must walk the bell‑line while the Mist‑Seer warns them not to follow the witchlights; the ferry token is their only safe exit. No further prep needed.

\subsection{The Travel Clock as Mood Tracker}
Do not just tick the clock on successes. Use it to track the \emph{feeling} of the journey:
\begin{itemize}
  \item \textbf{Segments 1–2: Optimism} – The road is good, the sun is out.
  \item \textbf{Segments 3–4: Fatigue} – Rations are boring. Minor arguments.
  \item \textbf{Segments 5–6: Desperation} – The rain hasn’t stopped for two days. Someone’s boot is leaking.
  \item \textbf{Segments 7–8: The Edge} – You’ll take any risk to be done.
\end{itemize}
When the clock fills, do not just say “you arrive.” Describe the stagger into town, mud‑caked and grateful.

\subsection{Genre Shifts Through Borders}
When the party crosses from one region to another, you are not just changing a map label – you are shifting the \textbf{dominant fantasy subgenre} (see \ref{subsec:travel-subgenres}). Use the travel deck to introduce the new region’s flavour immediately:
\begin{itemize}
  \item From \textbf{Ecktoria} (political intrigue) into \textbf{Mistlands} (folk horror): the road grows silent, fog rolls in, a bell tolls without wind.
  \item From \textbf{Acasia} (mercenary realism) into \textbf{Valewood} (fairy tale dark): the trees begin to whisper, and a path that wasn’t there appears.
\end{itemize}
The players should \emph{f e e l} the genre change before they see the border stone.

\subsection{Travelling Without a Map (No Hexes)}
Fate’s Edge does not use hex crawls. Travel is abstracted into \textbf{legs}, each with a Travel Clock (4, 6, 8, or 10 segments). You do not need to know how many miles it is from Silks trand to Ecktoria – you need to know how many \textbf{meaningful scenes} stand between them.  
This gives you complete freedom to insert or remove obstacles, side quests, and shortcuts without breaking verisimilitude.

% =========================================================
% EXISTING CLOCKS SECTION (expanded and clarified)
% =========================================================
\section{Clocks: Visible Pressure, Meaningful Choice}
\label{sec:clocks}
\index{Campaign Clocks}

Clocks are the primary way Fate’s Edge tracks narrative pressure over time.  
They are not secret timers or surprise punishments—they are \textbf{shared context} that tells the table what is at risk and what is coming.

A clock answers one of three questions:
\begin{itemize}
  \item \textbf{What is building?}
  \item \textbf{What happens if it completes?}
  \item \textbf{Who can affect it?}
\end{itemize}
If a clock does not clearly answer at least one of these, it is probably unnecessary.

\subsection{The Three-Clock Guideline (Per Scene)}
\label{subsec:three-clock-guideline}

As a rule of thumb, a scene rarely needs more than \textbf{three active clocks}:
\begin{itemize}
  \item \textbf{1 Immediate Clock} — What is about to go wrong \emph{right now} (e.g., \emph{Alarm Raised [4]}, \emph{Ritual Collapse [6]})
  \item \textbf{1 Situational Clock} — The pressure shaping player decisions (e.g., \emph{Crowd Turns Hostile [4]}, \emph{Structural Failure [6]})
  \item \textbf{1 Long-Term Clock} — How this scene echoes forward (e.g., \emph{Faction Retaliation [8]}, \emph{Corruption Advances [6]})
\end{itemize}
More clocks do not create more tension—they create \textbf{noise}. If players cannot name the clocks they are pushing against, they cannot make meaningful choices.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Concrete Example: The Heist at Lord Silas’s Manor]
  \textbf{Immediate Clock:} \emph{Guards Alerted [4]} – ticks each time the party makes noise or leaves evidence.\\
  \textbf{Situational Clock:} \emph{Patrol Pattern [6]} – ticks when the party delays; when full, a patrol arrives.\\
  \textbf{Long-Term Clock:} \emph{Silas’s Retaliation [8]} – ticks each time the party is identified; when full, Silas hires bounty hunters.
  
  \vspace{2mm}
  The players see all three clocks. They decide to rush the vault (risking \emph{Patrol Pattern}) rather than pick every lock carefully (which would tick \emph{Guards Alerted}). Their choice is informed and tense.
\end{tcolorbox}

\subsection{Fantasy Subgenres and the Travel Rules}
\label{subsec:travel-subgenres}
\index{regional subgenres}\index{travel rules!blending genres}

Fate’s Edge is built from distinct regions, each with its own dominant **fantasy subgenre**:
\begin{itemize}
  \item \textbf{Ecktoria} – Marble \& Fire (political intrigue, legal drama, fall of empire)
  \item \textbf{Acasia} – Broken Marches (low fantasy, mercenary realism, border feuds)
  \item \textbf{Mistlands} – Bells, Salt, and Breath (folk horror, isolated communities, supernatural dread)
  \item \textbf{Valewood} – Empire Under Leaves (fairy tale, forgotten magic, nature’s revenge)
  \item \textbf{Ykrul Steppe} – Storm on the Steppe (saga, nomadic honour, blood‑price)
  \item \textbf{Kahfagia} – Wakes and Storm‑Flags (nautical adventure, exploration, maritime trade)
  \item \textbf{Vhasia / Viterra} – Fractured Sun / Last Kingdom (courtly romance, chivalry, dynastic war)
\end{itemize}

When the party travels from one region to another, they cross between these subgenres. The \textbf{travel rules} (see Chapter 10 of the \textit{Player's Guide} and Section \ref{sec:gm-travel} above) are designed not merely to measure distance but to \textbf{blend these genres organically}. Use the travel deck and regional prompts to:
\begin{itemize}
  \item Introduce genre‑appropriate complications (a wraith‑touched fog in the Mistlands; a legal writ gone missing in Ecktoria).
  \item Allow the party to carry thematic elements from one region to the next (a curse from the Valewood might twist a courtly negotiation).
  \item Let the changing landscape reflect changing story stakes – travel is not filler but a \textbf{genre‑shifting engine}.
\end{itemize}

\subsection{Why Hidden Clocks Rarely Pay Off}
\label{subsec:hidden-clocks}
\index{clocks!hidden}

Hidden clocks often feel clever—but in practice, they weaken play.
\begin{itemize}
  \item Players cannot meaningfully react to pressure they cannot see.
  \item Surprise escalation feels arbitrary rather than earned.
  \item The GM ends up carrying tension alone instead of sharing it.
\end{itemize}
If a clock matters, \textbf{show it}. If it must remain unknown, foreshadow it through visible effects, rumors, or consequences—then reveal it as soon as the players can interact with it.

\textbf{Exception:} Short‑lived hidden clocks (e.g., \emph{Ambush Trigger [2]}) can work well; long‑term hidden clocks almost never do.

\subsection{Clocks Are Promises}
\label{subsec:clocks-promises}

Every clock you place on the table is a promise:
\begin{itemize}
  \item When it fills, \emph{something will happen}.
  \item That outcome will be \emph{consistent with what the players saw coming}.
  \item The players had \emph{opportunities to change the outcome}.
\end{itemize}
Clocks do not exist to punish failure—they exist to \textbf{make consequences legible}. When players choose to ignore a clock, that choice should feel informed, intentional, and costly.

\subsection{Mandate and Crisis (Campaign Clocks)}
\label{subsec:mandate-crisis}
\index{Campaign Clocks!Mandate}\index{Campaign Clocks!Crisis}

\textbf{Mandate (0–6)} represents the party’s public legitimacy and reputation. High Mandate means doors open; low Mandate means suspicion and obstacles.

\textbf{Crisis (0–6)} tracks the opposition’s rising pressure. High Crisis means enemies strike boldly and clocks tick faster.

Advance these clocks at the end of major scenes based on:
\begin{itemize}
  \item Clean loss or public cost paid → increase Crisis.
  \item Major victory or public recognition → increase Mandate.
  \item Asset neglect or scandal → affect both.
\end{itemize}

\subsection{Clock Evolution and Relationships}
\label{subsec:clock-evolution}

Campaign clocks should evolve as player actions change the world. Common progression:
\begin{enumerate}
  \item \textbf{Introduction} (0-2 segments): Threat becomes known.
  \item \textbf{Escalation} (3-4 segments): Threat gains momentum.
  \item \textbf{Crisis} (5-6 segments): Immediate danger to players/campaign.
  \item \textbf{Resolution} (7+ segments): Confrontation or transformation.
\end{enumerate}

Clocks can also relate to each other:
\begin{description}
  \item[Supporting] One clock's progress helps another (Plague Spread → Resource Scarcity).
  \item[Opposing] One clock's progress hinders another (Public Support ↓ Crime Rate ↑).
  \item[Cascade] One clock's resolution triggers another (War Ends → Reconstruction Begins).
\end{description}

% =========================================================
% THE CROWN SPREAD (now a major section, earlier and more prominent)
% =========================================================
\section{The Crown Spread: Seeding Your Campaign}
\label{sec:crown-spread}
\index{Crown Spread}\index{campaign design!oracular}

At \textbf{Session Zero}, draw the \textbf{Crown Spread}—a five-card ritual that seeds the campaign’s deepest themes, most dangerous rivals, and the very conditions of its ultimate resolution. This is not random chance but \textbf{oracular guidance} for the story to come.

\subsection{Drawing the Spread}
\label{subsec:crown-drawing}

Draw one card for each of the five fundamental aspects of your campaign:
\begin{itemize}
    \item \textbf{Spade} – The \textbf{Crown Site}: where destiny will be decided.
    \item \textbf{Heart} – The \textbf{Crown Rival}: who stands between the party and their goals.
    \item \textbf{Club} – The \textbf{Crown Pressure}: the relentless force that prevents complacency.
    \item \textbf{Diamond} – The \textbf{Crown Leverage}: the advantage that can turn the tide.
    \item \textbf{Wild} (draw a sixth card, or use the highest remaining card) – The \textbf{Hidden Force}: the unexpected element that will emerge when least expected.
\end{itemize}

\subsection{Interpreting the Spread}
\label{subsec:crown-interpreting}

\begin{fatebox}[Crown Spread Interpretation Guide]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Position} & \textbf{Examples and Questions} \\
\midrule
Spade (Site) & A ruined watchtower? A shrine built on forgotten truths? A battlefield where history repeats? \\
Heart (Rival) & A grieving widow with allies? A cult leader with apocalyptic visions? A spirit with ancient grievances? \\
Club (Pressure) & A spreading famine? An escalating curfew? A resource shortage that turns allies into competitors? \\
Diamond (Leverage) & A hidden letter exposing a noble’s treason? A seasonal endorsement from powerful factions? \\
Wild (Hidden) & A sympathetic deserter who knows secret passages? A Patron who steps from shadows? The site itself becomes a 10‑clock challenge? \\
\bottomrule
\end{tabularx}
\end{fatebox}

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=🌿 Worked Example: “The Bell at Blackwood Tower”]
  A GM draws:
  \begin{itemize}
    \item \textbf{7♠ (Spade)} → A ruined watchtower on a contested border, with a single bell that rings without wind.
    \item \textbf{Q♥ (Heart)} → A grieving widow who blames the party for her husband’s death.
    \item \textbf{9♣ (Club)} → A spreading famine; supplies dwindle, the militia conscripts farmers.
    \item \textbf{A♦ (Diamond)} → A hidden cache of pre‑Empire artifacts, including a sealed letter that exposes a noble’s treason.
    \item \textbf{K♠ (Wild)} → A deserter turned bandit who knows the watchtower’s secret passages and wants the artifacts for himself.
  \end{itemize}
  The GM now has: a location, a rival, a pressure, a secret advantage, and a wildcard. That is enough to run a 2‑3 session adventure without further prep.
\end{tcolorbox}

\subsection{Seasonal Evolution Framework}
\label{subsec:crown-evolution}

The Crown Spread elements should evolve through seasonal phases:
\begin{quote}
  \textbf{Winter (Establishment)}: Root themes take hold, initial conflicts emerge.\\
  \textbf{Spring (Growth)}: New elements sprout, alliances form, complications multiply.\\
  \textbf{Summer (Climax)}: Peak conflicts, major revelations, critical choices.\\
  \textbf{Autumn (Harvest)}: Consequences manifest, legacies established, new seeds planted.
\end{quote}

As the campaign progresses, deepen each element: add layers, introduce variations, connect to other Crown aspects, and evolve the stakes.

\subsection{Expanding Drawn Elements}
\label{subsec:crown-expanding}

When a Crown card’s theme becomes central to your campaign:
\begin{enumerate}
  \item \textbf{Deepen the Concept}: Add layers to the initial interpretation.
  \item \textbf{Introduce Variations}: Create related but distinct elements.
  \item \textbf{Connect to Other Elements}: Tie it to other Crown aspects.
  \item \textbf{Evolve the Stakes}: Raise the personal and cosmic implications.
\end{enumerate}

\section{Calling or Forcing the Crown}
\label{sec:calling-crown}
\index{Crown}\index{Finale}

The campaign reaches its crescendo when one of two thresholds is met—the moment when accumulated influence and mounting pressure collide in a final reckoning.

\begin{fatebox}[Finale Triggers and Conditions]
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Finale Type} & \textbf{Conditions and Narrative Implications} \\
\midrule
Player-Called & Mandate ≥ 6 and Crisis ≤ 3 – the party has earned the right to choose their moment of triumph. \\
Forced Finale & Crisis ≥ 6 regardless of Mandate – the world forces a confrontation that can no longer be avoided. \\
Balanced Finale & Both dials at 4–5 – a tense equilibrium where victory and defeat hang in perfect balance. \\
\bottomrule
\end{tabularx}
\end{fatebox}

\subsection{The Finale Procedure}
\label{subsec:finale-procedure}

When the Crown is called, run the three‑beat finale:
\begin{enumerate}
  \item \textbf{Reckoning}: Defend or sanctify the record of accomplishments. Draw upon the Rival’s established motives. Place the Pressure rail that will drive the scene forward.
  \item \textbf{Crossing}: Stage the kinetic rail (Escape/Hunt/Hazard) that threatens to end the scene prematurely if not managed carefully.
  \item \textbf{Coronation}: Use the Diamond Leverage to sign, seal, or swear the oath that cements the campaign’s legacy.
\end{enumerate}

\subsection{Twist Collision (Finale Clause)}
\label{subsec:twist-collision}

Exactly once, when the Rival’s Spade Twist contradicts their Club Belief, the table chooses:
\begin{itemize}
  \item GM gains +1 SB to complicate matters, or
  \item Players reduce two ticks total across the active rails, gaining breathing room.
\end{itemize}

\section{Legacy Conversion: Epilogue}
\label{sec:legacy-conversion}
\index{Legacy Conversion}\index{epilogue}

After the Finale, each PC draws 2 cards and answers epilogue prompts by suit. Then convert campaign elements into lasting legacy:
\begin{itemize}
  \item \textbf{Major Asset → Institution} (12 XP): A safehouse becomes a school, a spy ring becomes an intelligence service.
  \item \textbf{Seasonal Endorsement → Doctrine Rider} (4 XP): Temporary support becomes permanent policy.
  \item \textbf{Follower (Cap 3+) → Stationed NPC} (0 XP): Loyal companions become custodians of the new order.
  \item \textbf{Rival → Fixture}: Surviving adversaries become recurring elements of the setting’s fabric.
\end{itemize}

\subsection{Character Arc Management}
\label{subsec:character-arcs}

Help players develop meaningful character growth through the legacy process:
\begin{enumerate}
  \item \textbf{Establishment}: Define the character’s current state and potential conflicts.
  \item \textbf{Development}: Create opportunities for growth and choice.
  \item \textbf{Crisis}: Present challenges that test the character’s core beliefs.
  \item \textbf{Resolution}: Allow meaningful transformation based on choices.
\end{enumerate}

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=Arcs Generate Obligation and Patron Interest]
  \index{character arcs!obligation}\index{Patrons!in character arcs}
  A character’s personal arc should not exist in a vacuum. Tie it directly to the game’s resource systems:
  \begin{itemize}
    \item \textbf{Internal arcs} (fear, doubt, guilt) are perfect sources of \textit{Corruption} or \textit{Obligation}. A character struggling with guilt might be tempted by Malachai. A character seeking redemption might attract the attention of the Gallow’s Bell.
    \item \textbf{External arcs} (revenge, ambition, love) create \textit{Debt Clocks} with terrestrial patrons or generate \textit{Crisis} when they interfere with faction plans.
    \item \textbf{Whenever a player advances their arc, mark +1 Obligation to a relevant Patron or tick a Patron’s Interest clock.} The arc becomes a source of complications, not just backstory.
  \end{itemize}
  \textit{Example:} Sven wants to avenge his murdered brother. The GM notes that his brother was killed by a cult of Malachai. Each time Sven pursues a clue or confronts a cultist, the GM ticks \emph{Malachai’s Interest [6]}. When the clock fills, Malachai appears in a dream and offers Sven the power to find the killer – for a price.
\end{tcolorbox}

% =========================================================
% POSITION AS NARRATIVE LEVERAGE (shortened, cross‑referenced)
% =========================================================
\section{Position as Narrative Leverage}
\label{sec:position-narrative}
\index{Position!as narrative lever}

Position is the GM’s most powerful tool for calibrating narrative tension without altering core mechanics. The mechanical effect (re‑rolling one die) never changes, but its \emph{meaning} scales with the campaign’s scope:
\begin{itemize}
  \item \textbf{Tier I}: Position frames personal survival (“You escape with your life.”).
  \item \textbf{Tier III}: Position shapes regional outcomes (“The city’s fate hangs on your decision.”).
  \item \textbf{Tier V}: Position reshapes reality (“The universe holds its breath.”).
\end{itemize}

For full guidance on using Position, see Section \ref{sec:position-dv} in Chapter \ref{ch:gm-core}. 

\textbf{Remember:} Do not add mechanical bonuses to Position. Instead, deepen its narrative weight. A shift from Desperate to Controlled at Tier I means you live; at Tier V it means the world bends to your will.

% =========================================================
% BOON ECONOMY ACROSS TIERS (condensed)
% =========================================================
\section{Boon Economy Across Tiers}
\label{sec:boon-economy-tiers}
\index{boons!economy}\index{boons!tier scaling}

Boons are the players’ primary resource for re‑rolls, asset activations, and narrative leverage. Their availability changes dramatically as characters advance – and that shift is by design.

\begin{itemize}
  \item \textbf{Tier I–II}: Small dice pools (4–6 dice) → frequent Misses and Partials → 6–10 Boons per scene.
  \item \textbf{Tier IV–V}: Large dice pools (8–10 dice) → rare Misses, Partials common → 2–4 Boons per scene.
\end{itemize}

\textbf{Why this works:} Low‑tier characters need the cushion; high‑tier characters rely on skill, not luck. The scarcity at high tiers makes every Boon a strategic choice.

\subsection{Optional Rule: The Nine’s Boon}
\label{subsec:nines-boon}
\index{boons!optional rule}\index{dice!9 rule}

Some tables enjoy a little extra luck at any tier. This optional rule adds a small, controlled drip of Boons without breaking the high‑tier economy.

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=🎲 The Nine’s Boon (Optional), fonttitle=\bfseries]
  When you roll a natural \textbf{9} on any die (before re‑rolls), you gain \textbf{1 Boon}, provided you have not already earned the maximum of \textbf{2 Boons from failures this scene}. This Boon is generated immediately and may be spent normally.

  \medskip
  \noindent\textbf{Why 9?} The number nine is laden with threshold significance in Fate’s Edge – the un‑rung bell, the un‑poured cup, the step that is not counted. This rule leans into that theme without overlapping with the critical success of a 10.

  \medskip
  \noindent\textbf{Impact on high‑tier play:} Even with large dice pools, a 9 is a specific face, not a statistical flood. At 10 dice, you will see roughly one 9 per roll on average – still less Boon generation than low‑tier Misses. The rule adds a tiny safety valve without restoring the low‑tier abundance.

  \medskip
  \noindent\textbf{Note:} This Boon counts toward the per‑scene limit of 2 Boons from failures. If a player has already earned 2 Boons from Misses or Partials this scene, rolling a 9 grants nothing extra.
\end{tcolorbox}

% =========================================================
% RIVAL PARTIES (short version, cross‑referenced to full chapter)
% =========================================================
\section{Rival Parties}
\label{sec:rival-parties}
\index{rival parties}\index{GM tools!rivals}

A rival party is a group of NPC adventurers (or a faction cell) with goals that oppose, compete with, or complicate the PCs’ objectives. Rivals are not mere villains to be killed; they are \textbf{mirrors} and \textbf{pressure generators}.

For a full template and example (The Ashen Cord), see Section \ref{sec:rival-parties-full} in the GM’s Appendix. In brief:
\begin{itemize}
  \item Give them a name, composition (3–5 members), Patrons, and a visible \textbf{Rival Agenda clock} (6–8 segments).
  \item Tick their clock when the PCs ignore them or the GM spends SB on their off‑screen activities.
  \item When the clock fills, the rivals achieve a major milestone. Reveal their progress through rumors, discovered scenes, or confrontation.
  \item Use rivals to multiply Patron presence, create shifting alliances, and generate Story Beats from both sides.
\end{itemize}

% =========================================================
% THE CLOCKWORK ENGINE (short)
% =========================================================
\section{The Clockwork Engine: Tracking Tension}
\label{sec:clockwork-engine}
\index{clocks!introduction}

For detailed guidance on creating and managing tactical clocks, war clocks, and faction clocks, see the \textit{GM’s Guide} appendix on advanced clocks. The core principles remain:
\begin{itemize}
  \item Announce each clock and its triggers clearly.
  \item Use visible progression (tokens, whiteboard).
  \item Make consequences meaningful and fiction‑first.
\end{itemize}

\subsection{Advancement Rules}
Story Beats drive clock progression in measurable ways:
\begin{itemize}
  \item \textbf{1 SB}: Minor advancement (1 segment).
  \item \textbf{2-3 SB}: Moderate advancement (2 segments).
  \item \textbf{4+ SB}: Major advancement (3+ segments) or filling smaller clocks entirely.
\end{itemize}

\section{Between Sessions: The GM’s Sacred Trust}
\label{sec:between-sessions}
\index{GM responsibilities}

Between sessions, update clocks, track complication debt, and prepare scenes. Use the \textit{Between‑Session Checklist} in the GM’s Appendix. Remember: your preparation is the quiet magic that transforms random encounters into meaningful episodes.

\section{Campaign Pacing and Advanced Techniques}
\label{sec:campaign-pacing}

For guidance on session energy management, arc structure, scaling threats, branching narratives, and character arc planning, see the full \textit{Advanced GM Techniques} chapter. The principles are the same across all tiers: vary intensity, give arcs room to breathe, and let player choices drive consequences.

% =========================================================
% QUICK START CAMPAIGNS (condensed)
% =========================================================
\section{Quick Start Campaigns}
\label{sec:quick-start-campaigns}

\subsection{The Sunstone Gambit (Tier I–II, 4–6 sessions)}
A medium‑difficulty introduction to Fate’s Edge set in Acasia.  
\textbf{Key elements:} Ruined Sunstone Tower, rival treasure hunters, ancient guardian spirit, moral dilemma.  
\textbf{Session breakdown:} Commission → Travel → Exploration → Social maneuvering → Climax → Aftermath.

\subsection{Tier‑Appropriate Design}
\label{subsec:tier-design}
\begin{itemize}
  \item \textbf{Tier I–II (0–90 XP)}: Local scope, personal stakes, 3–5 session arcs.
  \item \textbf{Tier III (91–150 XP)}: Regional influence, branching arcs, 6–10 sessions.
  \item \textbf{Tiers IV–V (151+ XP)}: Continental scope, world‑shaping consequences, multi‑arc campaigns (15+ sessions).
\end{itemize}
For deeper guidance on high‑tier campaigns, see Chapter \ref{ch:gm-epic} (\textit{Tier IV and V: Running Epic Adventures}).

% =========================================================
% FINAL NARRATIVE REMINDER
% =========================================================
\section{Narrative First: The World Remembers}
\label{sec:narrative-first}
\index{narrative first}

Campaign design in Fate's Edge is not about railroading players along predetermined paths—it's about \textbf{responding to player choices} with consequences that accumulate like stones in a river, gradually shaping the flow of the narrative itself. Let the world shift in response to their actions. Let factions rise and fall based on their allegiances. Let the dice sing the song of a universe that reacts.

And when the Crown is finally crowned—when the last card is played and the final clock ticks to completion—let the echo of that moment be heard across the entire Amaranthine, a testament to stories well‑lived and consequences fully earned.

Remember: Your preparation between sessions is the quiet magic that transforms random encounters into meaningful episodes and mechanical challenges into memorable stories. The investment in this sacred trust pays dividends in player engagement, narrative coherence, and the creation of campaigns that will be remembered long after the final dice have been rolled.

\endinput