\chapter{Advanced GM Techniques}
\label{ch:gm-advanced}
\index{advanced techniques}

This chapter assumes you have mastered Fate’s Edge core rules. It offers tools for GMs running **experienced groups and higher‑tier campaigns (Tier III+)**. If you are new to the game, start with earlier chapters.

\section{Advanced Story Beat Management}
\index{Story Beats!advanced}

Story Beats (SB) are your primary tool for escalating tension. At high tiers, you will have more SB to spend – but also more narrative threads to manage.

\begin{tcolorbox}[
  colback=blue!5!white,
  colframe=blue!75!black,
  title=You Are Not Limited by SB,
  fonttitle=\bfseries
]
\index{Story Beats!as pacing tool}
\textbf{SB are a pacing tool, not a budget.} If the story needs a complication, introduce it. SB just tells you when players *want* escalation (by rolling 1s) and gives you a reservoir to draw from. Do not hoard SB; spend them freely to keep scenes dynamic.

\medskip
\noindent\textit{Example:} The party is interrogating a prisoner. Even without a roll, you sense the scene dragging. Spend 1 SB: “The prisoner’s eyes dart to the window – a signal. Someone is listening.” No roll, no budget – just pacing.
\end{tcolorbox}

\subsection*{Spending SB Across Multiple Threads}
Instead of pouring all SB into one complication, distribute them across several active clocks or fronts (see \ref{sec:clocks-fronts}). A scene might have:
\begin{itemize}
  \item \emph{Alarm Raised [4]} – advances from noise or failed stealth.
  \item \emph{Rival Closing In [6]} – advances from delays or loud magic.
  \item \emph{Patron Attention [8]} – advances each time a Rite is invoked.
\end{itemize}
Each Miss or Partial can tick one or two small clocks instead of a single big setback. This keeps pressure distributed and gives players multiple problems to solve.

\subsection*{Using SB to Advance Campaign Fronts}
When a scene directly relates to a long‑term front (e.g., a war, a spreading curse), spend SB to advance that front’s clock – even if the scene is otherwise a “success”. Example: The party defeats a cultist cell (clean success), but you spend 2 SB to tick \emph{Cult Influence [6]} – their network still grows elsewhere.

\subsection*{Banked SB from Player Complications}
Characters may start sessions with banked SB from unresolved complications (see \ref{sec:player-managed-modules-gm}). These count toward scene spending limits but should be spent on **thematic complications tied to that character’s background or Patron**, not generic setbacks. This makes personal storylines matter.

\subsection*{Example: Cascade Spend}
A single Miss generates 2 SB. The GM:
\begin{enumerate}
  \item Spends 1 SB to advance \emph{Alarm Raised [4]} by 1 (now 2/4 – guards are converging).
  \item Spends 1 SB to advance \emph{Rival Closing In [6]} by 1 (now 3/6 – rival scouts are spotted).
  \item Both clocks are now half full – the scene’s tension doubles without a single “big” complication.
\end{enumerate}

\section{Mastering the Deck of Consequences}
\index{Deck of Consequences!advanced}

The deck is not just a randomizer; it is a **thematic engine**. Use it to keep complications fresh and unpredictable, even at high tiers.

\subsection*{Shortcut Table: Suit + Rank → Typical Complication}
\label{subsec:deck-shortcut}
Use this table for fast interpretation at the table.

\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Suit} & \textbf{Low (2–5)} & \textbf{High (J–A)} \\
\midrule
Hearts (Emotion/Social) & Awkward social moment, minor suspicion & Betrayal, scandal, broken trust \\
Spades (Physical/Harm) & Bruise, small gear damage, stumble & Serious injury, structural collapse \\
Clubs (Resource/Cost) & Minor loss (rations, a few coins) & Major asset destroyed, debt called \\
Diamonds (Magic/Spiritual) & Omen, eerie noise, minor curse & Patron demand, reality bend \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Combining Draws (Suit Synthesis)}
Draw up to 3 cards per SB spend (minimum 1, maximum 3). Synthesise a single twist from the combination of suits, not just the highest rank.

\begin{itemize}
  \item \textbf{Hearts + Spades:} Emotional betrayal that leads to physical harm (“Your ally flees, and the portcullis drops on you.”)
  \item \textbf{Clubs + Diamonds:} Resource drain with magical backlash (“The ritual consumes your healing herbs and leaves a tainted scar.”)
  \item \textbf{Hearts + Clubs + Spades:} Social pressure plus resource loss plus injury – a crowd riots and loots your gear.
\end{itemize}

\subsection*{Using the Deck for Non‑Linear Scenes}
Draw a card to inspire flashbacks, montages, or off‑screen events. Example: During downtime, draw a Clubs card – “A fire at your safehouse.” Instead of playing it, you narrate a short scene of the party arriving to find the aftermath. The deck becomes a prompt generator, not merely a roll‑aftermath tool.

\subsection*{Custom Suit Tables}
For a specific campaign or region, replace the default suit meanings with custom ones. Example – Shadow Court campaign:
\begin{itemize}
  \item \textbf{Hearts:} Vows and obligations.
  \item \textbf{Spades:} Physical traps and barriers.
  \item \textbf{Clubs:} Political pressure and blackmail.
  \item \textbf{Diamonds:} Oaths and geasa.
\end{itemize}
This reflavours the deck without changing its structure.

\section{Advanced Travel and Genre Blending}
\index{travel!advanced}\index{regional subgenres}

Travel in Fate’s Edge is not filler – it is a way to **shift fantasy subgenres** as the party moves between regions. Use travel clocks (see Chapter 9 of the \textit{Player's Guide} and the separate \textbf{Travel Guide} document) and regional decks to create a journey that feels like turning pages from one story into another.

\subsection*{Designing Travel Legs as Mini‑Arcs}
Each travel leg can have its own dramatic arc: setup (departure), complication (en route), climax (arrival under pressure). Use a 4–6 segment clock for the leg. When it fills, they arrive – but the manner of arrival (e.g., chased, exhausted, bearing a new ally) changes the next scene.

\subsection*{Genre Shifting Example}
A journey from the \textbf{Mistlands (folk horror)} to \textbf{Ecktoria (political intrigue)} might have three legs:
\begin{enumerate}
  \item \textbf{Leg 1 (Mistlands):} Fog, bells, a wraith that speaks in riddles. (Horror)
  \item \textbf{Leg 2 (Borderlands):} A toll bridge, a corrupt clerk, a forged writ. (Low fantasy / crime)
  \item \textbf{Leg 3 (Ecktoria outskirts):} A noble’s hunting lodge, a dinner invitation, a whispered plot. (Intrigue)
\end{enumerate}
The travel deck for each leg should draw from the destination region’s themes – but the *complications* should reflect the transition.

\subsection*{Adding Rivals to Travel Legs}
Travel legs are excellent places to introduce rival parties (see \ref{sec:rival-parties}). A rival glimpsed at a waystation, tracks that cross yours, or a rumour that they passed through yesterday – all add pressure without direct confrontation.

\section{Faction Grand Strategy}
\index{factions!advanced}

At high tiers, factions act as “campaign engines.” Build them with **faction clocks** that advance whether the players intervene or not.

\subsection*{Faction Clock Examples}
\begin{itemize}
  \item \textbf{War [8]} – Advances when battles occur or supply lines are cut.
  \item \textbf{Influence [6]} – Advances when allies are won or propaganda spreads.
  \item \textbf{Intrigue [8]} – Advances when secrets are uncovered or betrayals happen.
\end{itemize}
These clocks are **visible to players** – they see the war brewing, even if they cannot stop it alone.

\subsection*{Player Actions and Faction Clocks}
When players act, they can tick a faction clock forward or backward. Example: Sabotaging a siege engine ticks \emph{War –1} (enemy’s clock) and ticks \emph{Their Own Mandate +1} (see \ref{subsec:mandate}). Document these changes openly. The world becomes a reactive board.

\subsection*{Running a “Faction Turn” Between Sessions}
\label{subsec:faction-turn}
For a highly political campaign, between sessions you can:
\begin{itemize}
  \item Ask players for their faction’s orders (e.g., “Send spies to the capital”).
  \item Roll a single die per order (Pass/Fail/Partial) using the faction’s relevant Asset (e.g., Spy Ring).
  \item Advance faction clocks accordingly.
  \item Start the next session with a brief “news from the front” segment.
\end{itemize}
This makes factions feel alive without heavy bookkeeping.

\subsection*{Example: Faction Turn with Rivals}
The rival party \textbf{The Ashen Cord} has their own faction clock: \emph{Cord’s Agenda [6]}. During the faction turn:
\begin{itemize}
  \item The GM rolls for the Cord’s order: “Secure the Sunstone fragment.”
  \item The roll is a Partial. The Cord finds the fragment but leaves evidence (tracks, a witness).
  \item The GM ticks \emph{Cord’s Agenda} by 1 and notes: “A merchant in Silkstrand saw a scarred woman matching Vesper’s description.”
  \item At the next session, the players hear the rumour. They now have a choice: pursue the Cord or stay on their own path.
\end{itemize}

\section{Designing High‑Tier Custom Content}
\index{custom content!high tier}

At Tiers IV–V, characters can reshape nations and realities. Custom content should grant **narrative permissions**, not merely bigger numbers.

\subsection*{Epic Talents (12+ XP)}
An epic talent might:
\begin{itemize}
  \item Allow rewriting one line of a Patron’s Rite (changing its cost or effect).
  \item Grant a once‑per‑session “truth edit” – the GM must accept one factual change to the scene.
  \item Let a character invoke a Patron’s name to retroactively avoid a TPK (once per campaign, with heavy cost).
\end{itemize}
These are not combat bonuses – they change what kind of story is being told.

\subsection*{Assets for Ruling Nations}
A Major Asset (12 XP) might be a kingdom. Its free use could be “order a levy” or “declare a holiday”. Its scene surge (1 Boon) could be “call in a noble favour” or “divert a river”. Such assets come with obvious upkeep (taxes, succession, rebellions) – and they can be threatened or lost.

\subsection*{Adventures for Demigods}
At the highest tiers, stakes should threaten **concepts**, not just lives. Examples:
\begin{itemize}
  \item A curse that erases the memory of a city.
  \item A war that will redraw the map for centuries.
  \item A Patron who will change the fundamental laws of magic.
\end{itemize}
Victory may not be guaranteed – but even failure should leave a permanent mark on the world.

\section{Running Mythic Scenarios (Tiers IV–V)}
\index{mythic scenarios}

At the apex of power, the game changes. Use these techniques to keep play grounded in choices, not points.

\subsection*{Handling Reality‑Altering Magic}
When a player attempts something cosmic (e.g., “I wish the volcano had never erupted”), do not say no. Instead, offer a **compromise**:
\begin{itemize}
  \item “You can do it, but it will cost you your connection to this place – you will never return.”
  \item “It works, but for the remainder of the session, the world is unstable – every roll generates an extra SB.”
  \item “You succeed, but your Patron demands a permanent sacrifice – one of your Talents is erased.”
\end{itemize}
The effect happens, but the price is felt.

\subsection*{Player Agency at Cosmic Scale}
\begin{tcolorbox}[
  colback=green!5!white,
  colframe=green!75!black,
  title=At Tier V, Every Choice May Erase an Alternative Timeline,
  fonttitle=\bfseries
]
\index{mythic scenarios!sacrifice}
Embrace this. Let players narrate what they are saving (or sacrificing) when they act.

\medskip
\noindent\textit{Example:} “I stop the ritual – but I let the rival escape.” That rival becomes a future problem. “I save the city – but I lose my connection to my Patron.” Now the player has a new arc: how to regain favor or find a new source of power.

\medskip
\noindent\textbf{Let the world break.} If the climax involves a collapsing reality, the battlefield itself can be a clock. Example: \emph{Reality Fracture [8]}. Each time a major event occurs (a spell, a death, an oath), tick the clock. When it fills, the scene shifts – time loops, gravity reverses, or the characters meet themselves from another timeline.
\end{tcolorbox}

\subsection*{Let the World Break (Clock Example)}
\index{clocks!reality fracture}
Set a visible clock: \textbf{Reality Fracture [8]}. Tick it when:
\begin{itemize}
  \item A character invokes a Patron’s name in desperation.
  \item A major magical effect is attempted.
  \item A character makes a life‑or‑death choice.
\end{itemize}
At 4 segments, strange echoes appear (ghostly duplicates, time slips). At 8 segments, the scene transforms: gravity inverts, doors lead to wrong locations, or the party faces an alternate version of themselves.

\section{Advanced Complex Scenarios}

\subsection*{Heists: Player‑Designed Plans}
Instead of planning the heist for them, ask each player: “How does your character contribute to the plan?” Build a clock (e.g., \emph{Heist Progress [8]}). Each player makes a roll using their chosen method (Stealth, Deception, Arcana, etc.). Success ticks the clock; failure adds complications (alarm, guard alerted). The GM spends SB from misses to introduce obstacles from the security system.

\subsection*{Mass Combat: Unit Clocks}
Instead of rolling for every soldier, treat opposing armies as a single \emph{War Clock [8]}. Player actions (command, sabotage, rally) tick the clock. When it fills, that side breaks or routs. Personal combat can happen in the same scene – but the mass conflict is abstracted to a single visible tracker.

\subsection*{Political Intrigue: Reputation as a Resource}
Track a \emph{Reputation Clock [6]} for the party in a given court. Social actions (gossip, favours, public deeds) tick it. When it fills, they become “trusted” – they can call in a major favour before the clock resets. But enemies also watch their rise, and may work to degrade the clock.

\section{When to Say Yes, No, and Yes, But – Advanced Rulings}
\index{GM rulings!advanced}

At high levels, players will attempt things that break normal bounds. Use these frameworks.

\subsection*{Compromise Solutions}
Instead of a flat “no”, offer a **scaled success** with a cost:
\begin{itemize}
  \item \textbf{Yes, but only for this scene.}
  \item \textbf{Yes, but you must mark a permanent Condition.}
  \item \textbf{Yes, but it will take longer – start a clock.}
  \item \textbf{Yes, but the GM gains 3 SB to spend on that effect.}
\end{itemize}
This keeps creativity alive while respecting the game’s tension economy.

\subsection*{Saying “No” Without Shutting Down}
If a plan is impossible (e.g., “I kill the god with a single arrow”), explain why: “The god is not physically present; you would need to unmake its last shrine first.” Then offer a path: “If you find the shrine and perform a ritual, you could banish it – you’ll need to gather three relics.” Turn the “no” into a new arc.

\subsection*{Let the Dice Decide – But Hedge}
For a truly impossible request, set an **extreme DV (6+)** and a **Desperate Position**. If they succeed, they get a lesser version of what they wanted, or a temporary victory. This respects the dice while maintaining verisimilitude.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=GM Golden Rule for Advanced Play]
\emph{The mechanics are not a ceiling – they are scaffolding. Build taller stories with them, but never let them crush creativity.}
\end{tcolorbox}