\chapter{Example Play and GM Survival Guide}
\index{example play}\index{gm survival guide}

This chapter provides two practical resources:
\begin{enumerate}
  \item A complete, annotated example session of \textbf{Fate’s Edge} (Tier I–II).
  \item A short \textbf{GM Survival Guide} – quick, actionable advice for common table situations.
\end{enumerate}
All other GM guidance (session zero, safety tools, scene framing, clocks, etc.) is covered in other chapters. This chapter focuses on \textbf{seeing the system in action} and \textbf{handling the unexpected} – while keeping you, the GM, as a fellow player at the table.

\section{The GM as Player: A Quick Refresher}
\index{GM as player}

Before diving into the example, remember: you are not a referee or a narrator robot. You are a \textbf{co‑creator of drama}, entitled to surprise, delight, and even a little mischief. Your tools (SB, clocks, Threat Pool) are not punishments – they are your way of playing the game. When you spend a Story Beat to tick a clock, you are making a move. When you describe a guard’s suspicious glance, you are adding tension. Enjoy it.

\section{Example Session: The Heist at Lord Silas’s Manor}
\index{example session}\index{heist}

This walkthrough follows three characters as they attempt to steal a sealed charter from a rival merchant prince. It demonstrates how scenes are framed, how pressure escalates, and how success, failure, Boons, and Story Beats work together – all while the GM plays an active, engaged role.

\subsection*{The Table}
\begin{itemize}
  \item \textbf{Valerius} — Ecktorian ex-legionary (Body 3, Warfare 2)
  \item \textbf{Elara} — Vhasian agent and negotiator (Wits 3, Sway 2)
  \item \textbf{Kael} — Dwarven geomancer (Lore 3, Geomancy 2)
  \item \textbf{GM} – Prepared to be surprised, to spend SB creatively, and to enjoy the ride.
\end{itemize}
The GM has prepared no fixed plot — only factions, motives, and clocks.

\subsection*{Campaign Context}
The group seeks a sealed charter proving their patron’s claim to a trade route. The charter is held by \textbf{Lord Silas}, a rival merchant prince.
\begin{itemize}
  \item Silas is politically insulated.
  \item His manor is guarded but discreet.
  \item Exposure is as dangerous as failure.
\end{itemize}

\subsection*{Scene 1: Information Gathering}
\textbf{GM frames:} “You are in Silkstrand’s lower markets at dusk. Silas’s manor rises above you. Guards are visible, but relaxed.” (The GM enjoys setting the atmospheric hook.)

\noindent\textbf{Stakes:} Success = identify an entry vector; Failure = guard suspicion escalates.
\noindent\textbf{Clock introduced:} \emph{Manor Alert} [4]

\paragraph{Action.} Elara scouts using \textbf{Wits + Subterfuge}. Position: Controlled.  
She rolls 2 Successes vs DV 3 → \textbf{Partial Success}.

\paragraph{Resolution.} GM rules:
\begin{itemize}
  \item Elara identifies a servant gate (progress achieved).
  \item GM advances \emph{Manor Alert} by 1 (cost).
  \item Elara gains \textbf{1 Boon}.
\end{itemize}
The fiction advances; pressure is visible. The GM is happy – a Partial is the most interesting outcome.

\subsection*{Scene 2: Entry}
The group scales the outer wall at night. Position: Desperate.  
\textbf{Stakes:} Success = inside undetected; Failure = immediate pursuit.

\paragraph{Action.} Valerius rolls \textbf{Body + Athletics}.  
He rolls 4 Successes including a 10 → \textbf{Clean Success with flourish}.

\paragraph{Resolution.}
\begin{itemize}
  \item The group enters cleanly.
  \item GM reduces future Position penalties inside.
  \item No Story Beat is spent.
\end{itemize}
Strong rolls relieve pressure without removing future danger. The GM celebrates the player’s cool moment.

\subsection*{Scene 3: Complication Inside}
Kael attunes to the stone to locate the vault. Position: Controlled.  
Kael rolls 0 Successes → \textbf{Miss}.

\paragraph{Resolution.}
\begin{itemize}
  \item Kael gains \textbf{2 Boons}.
  \item GM spends 1 SB to introduce a servant unexpectedly entering.
  \item GM advances \emph{Manor Alert} by 1.
\end{itemize}
The failure creates momentum rather than stalling play. The GM smiles – now they get to describe the servant’s wide eyes.

\subsection*{Scene 4: Social Pivot}
Elara calms the servant. Position: Risky.  
New clock: \emph{Servant Panic} [4].  
She rolls a \textbf{Success with Cost}.

\paragraph{Resolution.}
\begin{itemize}
  \item The servant is convinced.
  \item GM spends SB to advance \emph{Servant Panic} by 2.
  \item The group now operates under time pressure.
\end{itemize}
The GM is now thinking: “Will they make it before the servant screams?”

\subsection*{Scene Cut}
The GM cuts immediately: “You have the vault location. You also hear boots on marble.”  
This prevents overplaying and preserves tension. The GM enjoys the players’ anxious glances.

\subsection*{Scene 5: The Vault}
The vault is magically reinforced. Clock: \emph{Vault Integrity} [6].  
Kael attempts a geomantic bypass. He rolls Partial (2 Successes vs DV 4).

\paragraph{Resolution.} GM chooses proportional progress:
\begin{itemize}
  \item \emph{Vault Integrity} advances by 2.
  \item GM takes a Story Beat instead of increasing Alert.
  \item Kael gains 1 Boon.
\end{itemize}
Partials are \emph{GM-shaped outcomes}. The GM feels creative.

\subsection*{Scene 6: Extraction}
The vault opens as guards arrive. \emph{Manor Alert} at 3/4.  
Valerius creates a distraction while Elara escapes with the charter. The final roll succeeds — but the GM fills \emph{Manor Alert} as the cost.

\paragraph{Outcome.}
\begin{itemize}
  \item Charter secured.
  \item Silas knows he was robbed.
  \item New faction clock begins: \emph{Silas’s Retaliation} [6].
\end{itemize}
The GM is already planning the next session’s opening scene.

\subsection*{Session Close}
\noindent\textbf{End-of-session review:}
\begin{itemize}
  \item Boons generated primarily from Partials and Misses.
  \item Story Beats spent to escalate, not negate success.
  \item New clocks seeded future conflict.
\end{itemize}
The session ends with a clear win, lasting consequences, and visible momentum. The GM is satisfied – they were surprised, they played, and the table had fun.

\section{Example Session: The Rival’s Gambit}
\index{example session!rival party}\index{Ashen Cord}

This Tier I session introduces a \textbf{rival party} – the Ashen Cord – and demonstrates negotiation over combat, Boons spent on social actions, Patron intrusions, and failing forward.

\subsection*{The Table}
Same party: Valerius, Elara, Kael (now Tier I, 30–40 XP). The GM introduces a rival group: \textbf{The Ashen Cord} – Vesper (leader, shadow mage, serves Ikasha), Tallow (berserker, serves Malachai), Ember (cantor, serves Palinode), Cinder (scout, no Patron).

\subsection*{Campaign Context}
Both parties seek the same lost relic: a \textbf{Stone of Echoes} hidden in a ruined watchtower on the Acasian border. The relic is said to whisper secrets to those who attune to it.

\subsection*{Scene 1: The Approach}
GM frames: “The watchtower rises from the fog. You see another group approaching from the east – four figures, one carrying a wolf‑standard. They’ve seen you.”

\paragraph{Position:} Controlled. No clock yet – this is a social opening.

\paragraph{Action.} Elara steps forward to negotiate. “We don’t need to shed blood over this.” She spends \textbf{1 Boon} to improve her Position to Dominant before the roll.

\paragraph{Roll:} Presence + Sway (5 dice) vs DV 3. She rolls 3 Successes – Clean Success.

\paragraph{Resolution.} GM: “Vesper, the leader, agrees to parley. She offers a deal: ‘Whoever reaches the relic chamber first decides its fate. No backstabbing – word of honour.’” The rival party’s \emph{Cord’s Agenda [6]} clock is created but not advanced. The GM enjoys the tension of a fragile truce.

\subsection*{Scene 2: The Climb}
Both parties race up the tower’s crumbling stairs. The GM introduces a hazard: a collapsed section requiring an Athletics roll.

\paragraph{Action.} Valerius helps Kael across. Kael rolls Body + Athletics (4 dice) – Miss. Gains 2 Boons.

\paragraph{Resolution.} GM spends 1 SB: “Kael slips, but Valerius catches him. Kael’s pack tears open – a healing potion shatters. Mark 1 Supply segment lost.” The GM spends another 1 SB: “The noise alerts the Ashen Cord above – they gain a head start.” Advance \emph{Cord’s Agenda} by 1.

Failing forward: progress (they both cross) but costs and rival advantage.

\subsection*{Scene 3: The Relic Chamber}
Both parties arrive at the same time. The relic sits on a stone plinth. Vesper turns to the PCs: “Or we could settle this by rite – not blades. Each of us offers a secret. The one whose secret the stone accepts claims it.”

\paragraph{Social encounter.} The GM sets a \emph{Convince the Cord} [4] clock. Each successful social roll (Sway, Lore, or ritual) ticks the clock.

\paragraph{Action.} Elara tries to persuade Vesper to share first (Wits + Sway, DV 3). Partial (1 success vs DV 3). Gains 1 Boon. GM advances \emph{Cord’s Agenda} by 1 as Vesper’s patience thins.

\paragraph{Action.} Kael uses a geomantic reading (Spirit + Lore, DV 3). Clean Success. The GM ticks \emph{Convince the Cord} by 2. Kael’s Patron (The Sealed Gate) sends a small omen: a stone glows briefly.

\subsection*{Scene 4: Patron Intrusion}
Tallow (Malachai’s berserker) loses patience. He attacks the PCs – but Vesper shouts, “No!” The GM spends 2 SB: “Tallow’s axe swings wide, but Malachai’s mark on his hand flares. He stumbles back, cursing.”

\paragraph{Patron intrusion:} Malachai whispers to Tallow: \emph{“Break your word, and I’ll take your strength.”} Tallow is forced to abide by the truce.

\paragraph{Player response.} Elara spends 1 Boon to invoke her Bond with Kael (“We survived the heist together”) and gains +1 die on her next roll. She seals the deal: “We’ll share the stone’s first whisper – both parties benefit.”

\paragraph{Roll:} Presence + Sway (now +1 die = 6 dice) vs DV 3 – Clean Success. The \emph{Convince the Cord} clock fills.

\subsection*{Scene 5: Resolution}
Vesper agrees. Both parties touch the stone together. It whispers a location – a hidden vault beneath Silkstrand – and crumbles to dust.

\paragraph{Outcome.}
\begin{itemize}
  \item No combat. Boons spent on social actions, not blades.
  \item The Ashen Cord is not defeated but becomes a recurring rival (their \emph{Cord’s Agenda} is at 2/6). 
  \item A new clock: \emph{Silkstrand Vault Race} [6] – both parties will race to the vault next session.
\end{itemize}
The GM ends the session on a cliffhanger: “As you leave, you see Cinder (the Cord’s scout) slip away in the opposite direction – perhaps to beat you to the vault.”

\subsection*{Session Close Takeaways}
\begin{itemize}
  \item Negotiation over combat saved resources and created a more interesting relationship.
  \item Boons spent on social actions were as impactful as combat re‑rolls.
  \item Patron intrusions (Malachai) created tension without a fight.
  \item The rival party’s clock provides long-term pressure.
\end{itemize}

\section{High-Tier Play in Brief (Tier IV–V)}
\index{high-tier play}\index{boon scarcity}

At Tier IV–V, the core mechanics remain the same, but the context shifts. Here are the key differences, illustrated by a brief scenario.

\subsection*{What Changes}
\begin{itemize}
  \item \textbf{Larger dice pools} → more automatic successes, fewer Misses.
  \item \textbf{Fewer Boons earned} → Boons become precious, strategic currency.
  \item \textbf{Failure has wider consequences} – a missed roll can shift faction clocks or expose an asset.
  \item \textbf{Bonds and Assets} become the primary sources of Boons.
\end{itemize}

\subsection*{Example: House Morrenn Conflict}
A Tier IV party leads a regional faction against House Morrenn. Campaign clocks: \emph{Morrenn Ascendancy} [6] and \emph{Faction Stability} [6].

\paragraph{Scene: Council Confrontation.} Elara spends her only Boon on a Tier IV talent, succeeds, but the GM spends a Story Beat to mark her as a personal enemy – creating a new clock \emph{Targeted Retaliation} [4].

\paragraph{Scene: Asset Pressure.} Assassins strike the safehouse. No Boons remain, so Valerius accepts Harm and a worse Position to protect NPCs. \emph{Faction Stability} advances by 1 – people saw leadership under fire.

\paragraph{Scene: Risky Ritual.} Kael attempts a geomantic collapse to delay troops. He rolls a Miss, gains 2 Boons, but \emph{Morrenn Ascendancy} advances and long-term fallout (destabilised ley fault) appears.

\paragraph{Result.} The party ends with 1–2 Boons, strained assets, and multiple clocks near resolution. No DV inflation was needed – the stakes and resource scarcity did the work.

\subsection*{Key Takeaway}
At high tier, power comes from reach, not safety. Boons are for decisive moments, not routine re‑rolls. Failure is often chosen to regain leverage.

\section{GM Survival Guide}
\index{survival guide}\index{GM tips}

This section offers quick, actionable advice for the most common GM challenges. It assumes you have read the core rules and other chapters – this is the “cheat sheet” for the table.

\subsection*{When Players Go Off the Rails}
\begin{itemize}
  \item Do not panic. Ask: “What is your goal?” (not “How do you do that?”).
  \item Name the simplest obstacle in their way – then set Position and DV.
  \item If they ignore the obvious, let the world react. A missed roll means the guard they bypassed earlier shows up.
  \item Use \textbf{leading questions} to steer: “Are you trying to sneak past or create a distraction?” instead of “What do you do?”
  \item \textbf{GM as player:} Enjoy their creativity. Your job is to say “yes, but” and “no, but” – not to block them.
\end{itemize}

\subsection*{When Players Hoard Boons}
\index{boons!hoarding}
\begin{itemize}
  \item Remind them of the \textbf{Session 0 contract}: “Boons are not XP tokens – they are dramatic fuel. Spend them to shape the story.”
  \item Introduce \textbf{time pressure} that forces spending. “The ritual completes in three exchanges. You have 2 Boons – do you re‑roll that failed defense or save them?”
  \item Use the environment: a collapsing bridge, a spreading fire, a closing gate. “You can improve your Position to Dominant for 1 Boon – or risk a fall.”
  \item Offer a \textbf{Devil’s Bargain}: “Spend 1 Boon now to automatically succeed, but the rival will get a head start.”
  \item \textbf{GM as player:} Don’t punish hoarding – make spending tempting and visible.
\end{itemize}

\subsection*{How to Make Patrons Feel Present Without Overwhelming}
\index{Patrons!GM guidance}
\begin{itemize}
  \item Use \textbf{omens} before direct demands. A flickering candle, a dropped cup, a sudden chill.
  \item Send \textbf{dreams} or visions during rest. “You dream of a bell that tolls eight times – and then silence.”
  \item Have Patrons act through \textbf{small favors} that cost the players little but build relationship. “The Traveler clears the fog – but you feel a faint tug on your heel, a reminder that roads remember footsteps.”
  \item Escalate slowly: first a whisper, then a request, then a demand, then a reckoning.
  \item \textbf{GM as player:} Patrons are your co‑stars. Give them personality – a lisp, a favourite phrase, a sense of humour.
\end{itemize}

\subsection*{Handling Rival Parties}
\index{rival parties!GM guidance}
\begin{itemize}
  \item Keep rivals active \textbf{off-screen}. Tick their Agenda clock when the PCs dally or fail.
  \item Use \textbf{rumours} and \textbf{found scenes} to show their progress. A burned camp, a dead informant, a signature left behind.
  \item Bring them back when players least expect it – during a tense negotiation, at the climax of a heist, or as unexpected allies against a greater threat.
  \item Give rivals their own \textbf{Patrons} and \textbf{resources}. They should evolve just like PCs.
  \item \textbf{GM as player:} Rivals are your recurring NPC party. Enjoy running them – they are your chance to play, too.
\end{itemize}

\subsection*{When a Scene Stalls}
\begin{itemize}
  \item Ask a leading question: “Who are you trying to impress?” or “What are you afraid of here?”
  \item Tick a clock. “While you argue, the fire spreads – \emph{Collapse [4]} advances 1.”
  \item Cut to another player. Shift spotlight.
  \item \textbf{GM as player:} A stalled scene is your signal to introduce a new element. Have fun with it.
\end{itemize}

\subsection*{When You Don’t Know a Rule}
\begin{itemize}
  \item Make a ruling that keeps the story moving. Say: “We’ll do it this way for now; look it up later.”
  \item Default to DV 3, Controlled Position, and a consequence that feels right.
  \item Trust your gut. The system is robust.
  \item \textbf{GM as player:} Uncertainty is part of the game. Embrace it – the dice will surprise you.
\end{itemize}

\subsection*{How to Say “No” (Without Firing the Player)}
\begin{itemize}
  \item “No, but…” – offer a harder path, a cost, or a requirement.
  \item “Yes, but…” – let them succeed, then add a complication.
  \item “Let’s find out” – call for a roll with high DV and Desperate Position. The dice decide.
  \item \textbf{GM as player:} Saying “no” is boring. Saying “no, but” is a gift to the story.
\end{itemize}

\subsection*{Using Clocks Effectively}
\begin{itemize}
  \item Name clocks clearly: \emph{Alarm Raised [4]}, not “Bad Stuff [6]”.
  \item Only keep 1–3 clocks active per scene. More than that is noise.
  \item Tick a clock on a Miss or a Partial – even if the player otherwise succeeds.
  \item \textbf{GM as player:} Clocks are your visible toys. Tick them with relish.
\end{itemize}

\subsection*{Managing Player Spotlight}
\begin{itemize}
  \item Rotate by environment: scholar in the library, fighter on the wall.
  \item Use Bonds to pass focus: “You owe Rina – what do you do to protect her?”
  \item Quick check‑in: “While that happens, what is your character doing?”
  \item \textbf{GM as player:} Giving everyone a moment means you get to see different parts of the story.
\end{itemize}

\subsection*{When You Feel Burnt Out}
\begin{itemize}
  \item Call a break. Even five minutes resets the table.
  \item Ask players to recap – they will remind you of forgotten details.
  \item End the session early if needed. No game is better than a bad game.
  \item Use a published adventure or a random generator for one week. Give yourself a rest.
  \item \textbf{GM as player:} Your fun matters. If you are not enjoying it, talk to your group.
\end{itemize}

\subsection*{Final Survival Mantra}
\begin{quote}
\emph{Set the DV low. Make the world react honestly. Let the pressure do the work. And remember – you are playing too.} \\
— The Fate’s Edge GM Mantra
\end{quote}

If the table feels engaged, tense, and safe – you are running the game correctly. If you are also having fun, you are running it brilliantly.

\section{Session Zero Checklist (Handout)}
\index{Session Zero!checklist}
\label{sec:session-zero-checklist}

Use this checklist during your Session Zero. Write down the answers where everyone can see them.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Session Zero Checklist, fonttitle=\bfseries]
\begin{enumerate}
  \item \textbf{Lines \& Veils} – What content is off‑limits (Lines) or only implied (Veils)?
  \item \textbf{Tone} – Cinematic action? Grim survival? Political intrigue? Dark horror? Mix of several?
  \item \textbf{Campaign length} – One‑shot, short arc (3–6 sessions), or long campaign (12+ sessions)?
  \item \textbf{Character hooks} – Each player shares one goal and one fear for their character.
  \item \textbf{Patron interest} – Which Patrons (if any) are already watching the party?
  \item \textbf{Starting situation} – Where does the story begin? (Use the Crown Spread or a simple opener.)
  \item \textbf{House rules} – Any optional subsystems (Threat Pool, Deck of Consequences, etc.) you plan to use?
  \item \textbf{Scheduling \& attendance} – How often do you meet? What happens if a player misses a session?
  \item \textbf{Player-GM contract} – Remind the table: “We are co‑creators, not adversaries.”
  \item \textbf{Resource mindset} – “Boons are not XP tokens. Spend them. Embrace failure.”
\end{enumerate}
\end{tcolorbox}