\chapter{Advanced GM Techniques}
\label{ch:gm-advanced}
\index{advanced techniques}\index{GM!advanced}

This chapter assumes mastery of Fate's Edge core rules. It provides tools for GMs running **experienced groups and Tier III+ campaigns** where narrative complexity scales with player influence. New GMs should master core procedures first---these techniques build on that foundation.

\section{Advanced Story Beat Management}
\index{Story Beats!advanced}

Story Beats (SB) are your **narrative pulse**---not a budget to hoard, but a signal of where players *invite* escalation through their dice. At high tiers, you'll generate more SB (from larger dice pools), but your goal isn't to spend them all---it's to spend them *where they create the most compelling tension*. 

\subsection*{SB as a Pacing Tool, Not a Limit}
\begin{tcolorbox}[
  colback=blue!5!white,
  colframe=blue!75!black,
  title=You Are Not Limited by SB,
  fonttitle=\bfseries
,breakable]
\index{Story Beats!as pacing tool}
\textbf{SB measure player-generated narrative pressure, not your spending capacity.} If the fiction demands a complication, introduce it---even with 0 SB. SB simply tell you: *``Players rolled 1s here; they want the world to push back.''* Use them as fuel for existing threads, not as a constraint. 

\medskip
\noindent\textit{Example:} During a tense negotiation (no roll yet), you notice players leaning in. Spend 1 SB: ``The duke's envoy taps his signet ring---a signal. His hand drifts toward his sword hilt.'' No roll needed; you're responding to the *fiction's* tension, not chasing SB.
\end{tcolorbox}

\subsection*{Distribute SB Across Active Threads}
Instead of concentrating SB on one complication, **spend them to advance multiple visible timers or fronts** simultaneously. This creates layered tension where players must prioritize responses. 

\textbf{Effective distribution patterns:}
\begin{itemize}
  \item \textbf{Progressive tightening}: Spend 1 SB on \emph{Alarm Raised [4]} (now 2/4) and 1 SB on \emph{Rival Scouts [6]} (now 3/6). Both timers feel urgent but not overwhelming---players see converging threats.
  \item \textbf{Asymmetric pressure}: Spend 2 SB on \emph{Cult Ritual [8]} (now 6/8---imminent completion) and 1 SB on \emph{Innocent at Risk [4]} (now 2/4). Players face a classic ``save the many or stop the ritual'' dilemma.
  \item \textbf{Deception play}: Spend 1 SB on \emph{False Trail [4]} (now 3/4) to suggest a dead end, while secretly advancing \emph{Real Threat [6]} (now 2/6). Players waste resources on the wrong problem.
\end{itemize}
\textit{GM Tip:} Never let SB spending feel random. Each spend should answer: *``How does this complicate the player's declared goal or deepen an existing thread?''* 

\subsection*{Banked SB: Personalize the Escalation}
Characters may start sessions with banked SB from unresolved complications (e.g., a Debt Timer from character creation). **As GM, spend these banked SB exclusively on complications tied to that character's background, Patron, or unresolved debts**---never generic setbacks. This makes personal storylines matter mechanically.

\textbf{Example:} Lyra (Vilikari Rover) has 2 banked SB from her \textit{Two-Ledger Debt} complication. 
\begin{itemize}
  \item \textit{Wrong spend}: Using them to trigger a random bar fight. 
  \item \textit{Right spend}: 
    \begin{itemize}
      \item 1 SB: ``A Vilikari smuggler recognizes your ledger mark---he slides a note: *`The fence wants his silver tonight.'* (Advances \textit{Ledger Due [4]} timer +1)'' 
      \item 1 SB: ``As you leave, a courier bumps into you---your ledger is gone. Roll Wits + Subterfuge (DV 3) to notice before it's too late.'' 
    \end{itemize}
\end{itemize}
This turns banked SB into a direct callback to Lyra's story---no wasted escalation.

\section{Mastering the Deck of Consequences}
\index{Deck of Consequences!advanced}

The deck isn't a randomizer---it's a **thematic compass** for complications. Use it to ensure every twist feels inevitable in hindsight, even when surprising.

\subsection*{Shortcut Table: Suit + Rank $\rightarrow$ Narrative Core}
\label{subsec:deck-shortcut}
Keep this table visible for fast interpretation. Focus on the *narrative core*, not just mechanical effects.

\begin{table}[htbp]
\centering
\small
\begin{tabularx}{\textwidth}{lXX}
\toprule
\textbf{Suit} & \textbf{Low (2--5)} & \textbf{High (J--A)} \\
\midrule
Hearts (Emotion/Social) & Awkward social moment, minor suspicion & Betrayal, scandal, broken trust \\
Spades (Physical/Harm) & Bruise, small gear damage, stumble & Serious injury, structural collapse \\
Clubs (Resource/Cost) & Minor loss (rations, a few coins) & Major asset destroyed, debt called \\
Diamonds (Magic/Spiritual) & Omen, eerie noise, minor curse & Patron demand, reality bend \\
\bottomrule
\end{tabularx}
\caption{Shortcut table: Suit + Rank $\rightarrow$ Narrative Core}
\label{tab:deck-shortcut}
\end{table}

\subsection*{Synthesize, Don't Just List}
When drawing multiple cards (1 per SB spent, max 3), **force a single narrative synthesis**---not a list of effects. Ask: *``What one twist explains ALL these cards together?''*

\textbf{Synthesis examples:}
\begin{itemize}
  \item \textbf{Hearts + Spades}: Not ``betrayal + injury''---but *``Your lover shoves you aside as the trap triggers---you watch them take the dart meant for you.''* (Emotional betrayal \textit{causing} physical harm) 
  \item \textbf{Clubs + Diamonds}: Not ``resource loss + magic''---but *``The ritual consumes your last healing herb---but as the smoke clears, you see the sigil you drew is now bleeding ink.''* (Resource drain \textit{revealing} magical consequence) 
  \item \textbf{Hearts + Clubs + Spades}: Not ``social + resource + harm''---but *``The crowd's anger turns to looting---they smash your safehouse door while shouting *`You stole from the poor!'*''* (Social pressure \textit{driving} resource loss \textit{and} physical harm) 
\end{itemize}
\textit{GM Tip:} If you can't synthesize, redraw. A forced complication feels arbitrary; a synthesized one feels like fate.

\subsection*{Deck as a Prompt Generator}
Use the deck **beyond roll resolution**:
\begin{itemize}
  \item \textbf{Downtime worldbuilding}: Draw a card during downtime to inspire off-screen events. Example: Draw Clubs → ``A fire at your supplier's warehouse.'' Instead of resolving it, narrate: ``You arrive to find ashes and a note: *`Pay double or wait.'*'' 
  \item \textbf{Flashback seeds}: Before a scene, draw a card to frame a memory. Example: Draw Hearts → ``Your mentor's voice: *`Trust no one who smiles too wide*'.'' Players roleplay how this haunts their current choice. 
  \item \textbf{Montage triggers}: When skipping travel, draw a card to define the journey's tone. Example: Draw Spades → ``The road is littered with broken wagons---bandits have been active.'' 
\end{itemize}
\textit{This makes the deck a living part of your campaign's texture---not just a roll-aftermath tool.}

\subsection*{Custom Suit Tables for Campaign Flavor}
Refit the deck to your world's themes. **Replace suit meanings with campaign-specific tensions**---keep the structure, change the flavor.

\textbf{Example: The Crimson Throne Campaign (Patron-driven intrigue)}
\begin{itemize}
  \item \textbf{Hearts}: Vows and oaths broken (``Your sworn brother avoids your gaze'') 
  \item \textbf{Spades}: Physical manifestations of Patron displeasure (``The altar cracks as you speak---blood seeps from the stone'') 
  \item \textbf{Clubs}: Economic and political pressure (``The guildmaster slams the ledger: *`Your credit's revoked*''') 
  \item \textbf{Diamonds}: Divine omens and geasa (``A raven drops a black feather at your feet---*`The Watcher demands a life*''') 
\end{itemize}
\textit{Now every card draw screams ``Crimson Throne''---no explanation needed.}

\section{Adjudicating Thresholds with Narrative Weight}
\label{sec:thresholds-advanced}
\index{thresholds!adjudication}

Thresholds (doorways, crossroads, riverbanks, sleep/waking edges) are where the world's memory is thinnest. **As GM, your job isn't to apply rigid effects---it's to use the threshold as a narrative lever** that ties player action to consequence.

\subsection*{Thresholds as Conversation Starters}
When a player says *``I cross the threshold,''* **respond with a question that demands narrative engagement**, not a die roll. Use the table below as a guide for *what to ask*---not as a mechanical lookup.

\begin{table}[htbp]
\centering
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Threshold} & \textbf{Your Narrative Prompt (Ask the Player)} \\
\midrule
Doorway (unwarded) & ``What do you *feel* as you step through? A chill? A whisper? A sudden certainty?'' \\
\midrule
Doorway (warded) & ``What offering do you leave? A coin? A promise? Or do you rush through, ignoring the sigils?'' \\
\midrule
Crossroads & ``Which path feels *right*---not safe, not easy, but *true* to who you are now?'' \\
\midrule
Riverbank & ``What do you *see* in the water's face? Your past? A warning? A face you hoped to forget?'' \\
\midrule
Bridge & ``What memory does this bridge *call up* for you? A betrayal? A promise kept? A life left behind?'' \\
\midrule
Sleep/waking edge & ``What dream clings to you as you wake? Or what fear keeps you from sleeping?'' \\
\bottomrule
\end{tabularx}
\caption{Thresholds as narrative prompts (not mechanics tables)}
\label{tab:threshold-prompts}
\end{table}

\subsection*{Threshold Timers for Liminal Locations}
For places thick with threshold energy (haunted mansions, fey crossroads, battlefields), start a \textbf{Threshold Instability [4--6]} timer. \textbf{Tick it when players cross thresholds without narrative respect}---not just mechanically, but when they *ignore the fiction's weight*.

\textbf{What counts as ``disrespect'' (tick the timer):}
\begin{itemize}
  \item Crossing a doorway without acknowledging its spirit (``You shove the door open---no pause, no word'') 
  \item Treating a crossroads as just a turn (``You take the left path without noticing the three stones arranged in a triangle'') 
  \item Walking a riverbank while eyes are glued to a map (``You don't see the heron watching you---it flies off silently'') 
\end{itemize}
\textit{When the timer fills, the threshold inverts}: 
\begin{itemize}
  \item Doors open to wrong rooms (``You enter the pantry---find yourself in the lord's bedchamber'') 
  \item Crossroads show false paths (``The left road now leads *back* to where you started'') 
  \item Riverbanks reveal drowned truths (``The water parts---you see your own face, eyes closed, sinking'') 
  \item Bridges demand memories crossed (``To cross, you must speak aloud the name of someone you've failed'') 
  \item Sleep/waking edges blur (``You wake---but is this dawn, or the moment you died?'') 
\end{itemize}
\textit{GM Tip:} Describe the inversion through the player's senses---don't just state it. Make them *feel* the world's memory shifting.

\subsection*{Player-Driven Threshold Depth}
When players specify *how* they cross a threshold (**``I open the door with my left hand''** or **``I pour wine on the threshold before stepping over''**), **use that fiction to set Position and Effect**. 

\textbf{Effect thresholds:}
\begin{itemize}
  \item \textbf{Respectful crossing} (removing boots, libation, whispered name): Start \textbf{Controlled} or grant \textbf{+1 Effect} (e.g., ward doesn't trigger, spirits acknowledge you) 
  \item \textbf{Brash crossing} (kicking doors, ignoring omens, rushing): Start \textbf{Desperate} or impose \textbf{--1 Effect} (e.g., ward triggers, spirits are offended) 
  \item \textbf{Threshold-specific action} (e.g., ``I step *over* the river stone, not on it''): Grant a situational bonus (e.g., +1 die to avoid water spirits) 
\end{itemize}
\textit{Example:} A player says, ``I back away from the altar, eyes down, murmuring my vow.'' As GM: ``Your humility is seen---the altar's glow softens. Start \textbf{Dominant} for this interaction, and gain +1 Effect on your next Persuasion roll with the temple priests.''

\section{Faction Grand Strategy: Making the World Breathe}
\label{sec:faction-grand-strategy}
\index{factions!grand strategy}\index{factions!advanced}

At Tier III+, factions aren't static backdrops---they're **campaign engines** with their own goals, timers, and reactions to player actions. Your job isn't to simulate every faction member---it's to make factions *feel* alive through visible timers, meaningful player agency, and emergent consequences.

\subsection*{Factions as Visible Timers with Personality}
Represent each major faction with **1--3 visible timers** that reflect their core drives. Make timers *narrative*, not abstract:

\textbf{Example Faction Timers:}
\begin{itemize}
  \item \textbf{The Iron League (Mercenary Coalition)}: 
        \begin{itemize}
          \item \emph{Paymaster's Patience [6]}: Advances when players delay payment or take jobs without consulting the League. When full: ``The League recalls all contracts---your mercenary contacts refuse jobs until you clear arrears.'' 
          \item \emph{Recruitment Drive [4]}: Advances when players showcase skill or generosity. When full: ``The League sends a captain---*`We've heard of your deeds. Join us, and we'll make you legend*'.'' 
          \item \emph{Internal Rift [8]}: Advances when players exploit League divisions or fail to defend allies. When full: ``Two captains draw swords at the muster---*`You chose this weakness*'.'' (Players must pick a side or watch the League fracture) 
        \end{itemize}
  \item \textbf{The Veiled Cabal (Spy Network)}: 
        \begin{itemize}
          \item \emph{Mislead the Hunt [6]}: Advances when players spread false intel or create diversions. When full: ``Your false tip bears fruit---the enemy marches into an ambush. A coded note arrives: *`Well played*'.'' 
          \item \textbf{Assets Exposed [8]}: Advances when players use Veiled Cabal assets carelessly (e.g., burning a safehouse). When full: ``A safehouse door is kicked in---*`We know your methods*'.'' (Players lose access until they rebuild trust) 
          \item \textbf{Whisper Network [4]}: Advances when players gather secrets discreetly. When full: ``A child presses a note into your hand: *`The duke meets the assassin at midnight*'.'' 
        \end{itemize}
\end{itemize}
\textit{GM Tip:} Name timers after *faction motivations* (not mechanics). ``\emph{Paymaster's Patience}'' tells a story; ``\emph{Timer 3}'' does not.

\subsection*{Player Actions as Faction Diplomacy**
When players act toward a faction, **adjudicate the outcome through the lens of the faction's timers and values**---not just a die roll. 

\textbf{Effective adjudication flow:}
\begin{enumerate}
  \item \textbf{Hear the player's intent}: ``I want the Iron League to fund our expedition.'' 
  \item \textbf{Ask: What does this action *say* to the faction?}: 
        \begin{itemize}
          \item *Respectful*: ``I've cleared their name in three towns---I bring honor.'' (Aligns with \emph{Recruitment Drive}) 
          \item *Transactional*: ``I'll give you 20% of any loot we find." (Neutral---may advance/decline timers based on context) 
          \item *Exploitative*: ''I'll take your funds, but I won't share credit.`` (Risks \emph{Internal Rift}) 
          \item *Threatening*: ''Fund us, or we'll tell everyone you take bribes from slavers.`` (Risks \emph{Paymaster's Patience} + \emph{Internal Rift}) 
        \end{itemize}
  \item \textbf{Set DV based on faction's current state}: 
        \begin{itemize}
          \item If \emph{Paymaster's Patience} is near full: Higher DV (they're wary of being used) 
          \item If \emph{Recruitment Drive} is active: Lower DV (they're eager for allies) 
          \item If \emph{Internal Rift} is ticking: DV varies wildly by who you're speaking to 
        \end{itemize}
  \item \textbf{Apply Outcome Matrix, then spend SB to show faction reaction}: 
        \begin{itemize}
          \item \textbf{Clean Success}: Faction reacts as hoped---but spend 1 SB to show a hidden cost (''The captain nods, but his eyes flick to your rival's insignia---*`We'll see how long this lasts*'.``) 
          \item \textbf{Success with SB}: Faction agrees, but SB shows strings attached (''They fund you---but the note says *`Return with the artifact, or we reclaim it*'.``) 
          \item \textbf{Partial}: Faction hesitates---but spend SB to show what would tip them (''The lieutenant frowns: *`Bring us proof of the route's safety, and we'll talk*'.``) 
          \item \textbf{Miss}: Faction rejects---but spend SB to show *why* and what they'll do next (''The master slams the table: *`Your deeds are lies*'.`` Then advance \emph{Paymaster's Patience} by 1---they're now less trusting of *all* outsiders) 
        \end{itemize}
\end{enumerate}
\textit{This turns faction interaction into roleplay---where dice resolve *feasibility*, but fiction resolves *meaning*.}

\subsection*{Running the Faction Turn Between Sessions**
For politically dense campaigns, **use the faction turn to make the world evolve *while players rest***---without heavy bookkeeping. 

\textbf{GM Procedure:}
\begin{enumerate}
  \item \textbf{Between sessions, email players}: ''The Iron League has issued orders. Reply with how your character responds *in character* (e.g., *`I send 50% of my loot as tribute'* or *`I ignore them---I've got a score to settle*'). Deadline: [time]." 
  \item \textbf{As GM, resolve orders using faction assets}: 
        \begin{itemize}
          \item For each order, roll using the faction's relevant Asset (e.g., Spy Ring for intel missions, Treasury for funding requests). 
          \item \textbf{Outcome Matrix applies}: 
                \begin{itemize}
                  \item Clean Success: Order succeeds flawlessly 
                  \item Success with SB: Order succeeds but with complication (note SB for GM) 
                  \item Partial: Order partially succeeds---note what failed 
                  \item Miss: Order fails---note why and what faction does instead 
                \end{itemize}
          \item \textbf{Translate outcome to narrative}: 
                \begin{itemize}
                  \item *Order: ``Send spies to the capital''* 
                  \item *Roll: Partial (Stealth + Subterfuge = 4 successes vs DV 5)* 
                  \item *GM spends 1 SB (from roll)* 
                  \item *Result*: ``Spies reach the capital gates---but are turned away. They report: *`The gates are sealed. Word is, the League suspects a leak*'.'' (Advances \emph{Mislead the Hunt} timer +1 via spent SB; notes that spies failed to enter) 
                \end{itemize}
        \end{itemize}
  \item \textbf{Start next session with ``News from the Front''}: 
        \begin{itemize}
          \item Spend 2--3 minutes describing faction outcomes *as in-world rumors, reports, or events*. 
          \item \textit{Example:} ``As you break camp, a League rider approaches---dusty, wary. *`The Paymaster sends word: your tribute's accepted. But he asks---why did your scout ask about the north gate?*''' (Players now know their Partial success created suspicion) 
          \item *Then*: ``Meanwhile, a Veiled Cabal agent slips you a folded silk---*`The duke's guard changes at 2 AM. The silk's dyed with midnight blossom*'.'' (Reveals a Whisper Network success from their faction turn) 
        \end{itemize}
\end{enumerate}
\textit{Why this works:} Players see their choices *matter in the fiction*---not just as XP gains. The faction turn makes the world reactive *between* sessions, so every session starts with consequences to engage with.

\subsection*{Faction Timers as Complication Engines**
Neglecting a faction (ignoring their timers, acting against their values) **should advance their timers and create complications**---not be invisible. 

\textbf{Consequence ladder for faction neglect:}
\begin{itemize}
  \item \textbf{1 session neglected}: Timer advances normally; spend 1 SB for a minor friction (``The League envoy's smile is tight---*`We'll discuss your conduct at muster*'.'') 
  \item \textbf{2 sessions neglected}: Timer advances +1 extra; spend 2 SB for a meaningful setback (``The League recalls 50% of your funding---*`Until you prove your loyalty*'." Players lose access to a Major Asset benefit until they mend ties) 
  \item \textbf{3+ sessions neglected}: Timer fills; trigger a faction-specific consequence (''The Iron League assemblies a vote---*`Expel the oathbreakers*'.`` Players are now *hostile* to the League until they complete a major service quest) 
\end{itemize}
\textit{GM Tip:} Make neglect escalate *in faction-specific ways*. Ignoring the Veiled Cabal shouldn't just give --1 die---it should make their safehouses burn, their informants vanish, or their omens turn hostile.

\section{Advanced Timer Design: Creating Self-Sustaining Tension}
\label{sec:advanced-timers}
\index{timers!advanced}\index{timers!interacting}\index{timers!cascading}

Timers become narrative powerhouses when they **interact, cascade, or hide consequences behind success**. Stop tracking isolated timers---design them as parts of a living system.

\subsection*{Interacting Timers: Timers That Talk to Each Other**
Two timers where advancing one *automatically* affects the other (no player action needed). Use these to model systemic pressures.

\textbf{Patterns for interacting timers:}
\begin{itemize}
  \item \textbf{See-saw}: Advancing Timer A *retreats* Timer B (and vice versa). 
        \textit{Example:} \emph{City Trust [6]} and \emph{Guard Vigilance [6]}. 
        \begin{itemize}
          \item Players spend time in taverns, boosting morale → \emph{City Trust} advances → \emph{Guard Vigilance} retreats 1 (guards are distracted by festivities) 
          \item Players crack down on crime → \emph{Guard Vigilance} advances → \emph{City Trust} retreats 1 (citizens feel oppressed) 
        \end{itemize}
        \textit{GM Tip:} When you advance one timer, *immediately* check if its partner should retreat. Describe the shift: ''The market's buzz is louder tonight---but the watch stands stiff-eyed at every corner.`` 
  \item \textbf{Threshold gating}: Timer A only advances *when* Timer B is in a specific state. 
        \textit{Example:} \emph{Plague Spread [8]} only advances if \emph{Sanitation Efforts [4]} is below 3 segments. 
        \begin{itemize}
          \item If sanitation is good (Timer ≥3), plague spreads slowly 
          \item If sanitation fails (Timer <3), plague spreads fast 
        \end{itemize}
        \textit{GM Tip:} When players act on sanitation, say: ''You scrub the wells---but the foul smell persists. The *Sanitation Efforts* timer retreats\dots but the *Plague Spread* timer doesn't advance yet. You need to do *more*.`` 
\end{itemize}

\subsection*{Cascading Timers: Dominoes of Escalation**
When one timer fills, it *creates* a new timer---or advances an existing one in a meaningful way. Models how small problems snowball.

\textbf{Patterns for cascading timers:}
\begin{itemize}
  \item \textbf{Threshold breach}: Timer fills → new timer appears representing irreversible change. 
        \textit{Example:} \emph{Dam Integrity [6]} fills → \emph{Flood Warning [4]} appears. 
        \begin{itemize}
          \item While dam holds: Players can reinforce it (advance \emph{Integrity}) or ignore it (risk timer) 
          \item When dam breaks: \emph{Flood Warning} appears---now players must evacuate, divert water, or face consequences 
          \item \textit{GM Tip:} When the dam breaks, don't just say ''It broke.`` Describe: ''The groan shakes your bones---a jagged tear splits the stone. Water isn't leaking; it's *pouring*.`` Then introduce the new timer. 
        \end{itemize}
  \item \textbf{Escalating threat}: Timer fills → existing timer advances *faster*. 
        \textit{Example:} \emph{Spy Network [6]} fills → \emph{Counterintelligence [4]} now advances 2 sections per tick (instead of 1). 
        \begin{itemize}
          \item Early game: Spies gain intel slowly; counterintelligence lags 
          \item Late game: Spies are effective---but every success triggers a rapid counter-response 
        \end{itemize}
        \textit{GM Tip:} Announce the shift narratively: ''The captain slams his fist: *`They've cracked our code*'.`` Then update the timer rule. 
  \item \textbf{Resource depletion}: Timer fills → related timer begins *counting down*. 
        \textit{Example:} \emph{Grain Stores [8]} fills → \emph{Famine Clock [6]} starts counting *down* from 6 to 0. 
        \begin{itemize}
          \item While stores full: Players can trade, feast, or invest 
          \item When stores empty: \emph{Famine Clock} ticks down---each segment is a week until starvation hits 
          \item \textit{GM Tip:} When the clock hits 3, describe: ''The bread is sawdust. Children's bellies swell with hunger. You have *three weeks* to find food---or watch them waste away.`` 
        \end{itemize}
\end{itemize}

\subsection*{Hidden Consequences Behind Success**
Even a Clean Success can advance a long-term timer---**but only if the fiction supports it**. Never hide consequences arbitrarily; make them feel inevitable.

\textbf{When to hide consequences behind success:}
\begin{itemize}
  \item \textbf{The success creates the problem}: 
        \textit{Example:} Players assassinate a tyrant (Clean Success). Off-screen, \emph{Martyr Cult [8]} advances +2---the death makes him a symbol, radicalizing followers. 
        \textit{GM narration}: ''The blade falls---but as the crowd cheers, a child presses a flower into the tyrant's hand. *`He died for us*'.`` 
  \item \textbf{The success requires a hidden cost}: 
        \textit{Example:} Players negotiate a trade pact (Clean Success). Off-screen, \emph{Tribute Debt [6]} advances +1---the pact requires annual payments they can't afford. 
        \textit{GM narration}: ''The seals are stamped---but as you leave, the envoy slips you a clay tablet: *`Year one's tribute is due at harvest*'.`` 
  \item \textbf{The success attracts unwanted attention}: 
        \textit{Example:} Players discover a cure for a plague (Clean Success). Off-screen, \emph{Patron Ire [6]} advances +1---the miracle drew a jealous Patron's gaze. 
        \textit{GM narration}: ''The vial glows---but as you cap it, the scent of ozone draws a shadow in the corner. *`Such power should not be wielded by mortals*'.`` 
\end{itemize}
\textit{GM Tip:} When hiding consequences, **always describe the narrative trigger**. Players should think: *''Oh---that makes sense.``* Not: *''Wait, where did that come from?``*

\section{Designing High-Tier Custom Content: Permissions, Not Power}
\label{sec:custom-content-high-tier}
\index{custom content!high tier}

At Tiers IV--V, custom content shouldn't just add bigger numbers---it should grant **narrative permissions** that let players reshape the story's boundaries. These are *story tools*, not combat bonuses.

\subsection*{Epic Talents (12+ XP): Rewriting the Rules of Engagement**
An epic talent alters *what kind of story is being told*, not just damage output. 

\textbf{Examples (with GM guidance):}
\begin{itemize}
  \item \textbf{Rite Reweaver}: Once per session, rewrite *one line* of a Patron's Rite you know (changing its cost, effect, or Obligation). 
        \begin{itemize}
          \item \textit{Player says}: ''I invoke the Oath of Flame \& Light's Rite of Lay on Hands---but I change *`reduce Harm by 1'* to *`remove 1 Harm and 1 Fatigue*'.`` 
          \item \textit{GM adjudicates}: ''As you speak the Rite, the words feel like clay in your mouth. You reshape it---but the light flickers. *`The oath is bent, not broken*'.`` (Advances Patron's Obligation +1 for the tampering) 
          \item \textit{Why it works}: It's not ''more healing``---it's *''I can make the Patron's power serve my need, but at a cost to our bond.``* 
        \end{itemize}
  \item \textbf{Truth Edit}: Once per session, declare one factual change to the current scene---the GM *must* accept it. 
        \begin{itemize}
          \item \textit{Player says}: ''The assassin's dagger---it's not poisoned. It's just a knife.`` 
          \item \textit{GM must accept**: Describe the dagger as mundane, then spend SB to show why the lie mattered (''The wound is clean---but as you pull it free, you see the hilt bears the sigil of the Black Ledger. *`They wanted it to look like an accident*'.``) 
          \item \textit{Why it works}: It's not ''I win the roll``---it's *''I can reshape perception, but the truth leaks out in other ways.``* 
        \end{itemize}
  \item \textbf{Patron's Name (Once/Campaign)**: Invoke a Patron's name to retroactively avoid a TPK---once per campaign, with heavy cost. 
        \begin{itemize}
          \item \textit{Player says}: ''I cry out *`Oath of Flame \& Light!*' as the dragon's fire hits us.`` 
          \item \textit{GM adjudicates**: The fire stops---but you mark 3 Harm (Level 3, incapacitated) and lose all connection to this Patron forever. Describe: ''The flames halt mid-air---but as they vanish, you feel the bond snap like a rotten rope. Your vision grays; you know you'll never feel their warmth again.`` 
          \item \textit{Why it works}: It's not ''I survive``---it's *''I can call on a Patron's ultimate power\dots but it severs the tie that made it meaningful.``* 
        \end{itemize}
\end{itemize}
\textit{Core principle:} Epic talents should make players ask: *''What story am I *allowed* to tell now---and what do I sacrifice to tell it?``*

\subsection*{Assets for Ruling Nations: Power with Strings Attached**
A Major Asset (12 XP) representing national control should come with **inherent, unavoidable complications** that scale with use.

\textbf{Examples (with GM guidance):}
\begin{itemize}
  \item \textbf{Kingdom Charter}: Free use: ''Order a levy`` (raise troops/taxes). Scene Surge (1 Boon): ''Divert a river`` (flood enemy lines, irrigate fields). 
        \begin{itemize}
          \item \textit{Inherent complications}: 
                \begin{itemize}
                  \item Levy order advances \emph{Public Resentment [6]} (more troops = more anger) 
                  \item River diversion advances \emph{Ecological Backlash [6]} (floods breed disease; droughts kill crops) 
                  \item Using either Surge more than once/session advances \emph{Tyranny Flag [4]} (players are seen as oppressive) 
                \end{itemize}
          \item \textit{GM narration when using Free use}: ''The conscript lines form---but as you pass, you hear a mother whisper: *`They took my son for the war*'.`` 
          \item \textit{GM narration when using Scene Surge**: ''The river bends to your will---but downstream, the fish float belly-up. An old fisherman crosses himself: *`The river god is angry*'.`` 
        \end{itemize}
  \item \textbf{Imperial Archive}: Free use: ''Gain a Boon on a Lore roll.`` Scene Surge (1 Boon): ''Convene a symposium of scholars.`` 
        \begin{itemize}
          \item \textit{Inherent complications}: 
                \begin{itemize}
                  \item Free use advances \emph{Forbidden Knowledge [6]} (the more you learn, the more you uncover secrets that endanger you) 
                  \item Scene Surge advances \emph{Scholarly Factions [4]} (funding the symposium makes enemies among traditionalists) 
                \end{itemize}
          \item \textit{GM narration when using Free use**: ''You trace the citation---but as you close the tome, a cold spot blooms on your neck. *`Some knowledge leaves a mark*'.`` 
          \item \textit{GM narration when using Scene Surge**: ''The scholars debate---but as they leave, you see two groups glaring at each other over coffee. One mutters: *`They'll burn the archives for this*'.`` 
        \end{itemize}
\end{itemize}
\textit{Core principle:} National power isn't a bonus---it's a *web of obligations*. The Asset's value isn't in what it *does*, but in what it *demands*.

\subsection*{Adventures for Demigods: Stakes That Reshape Reality**
At the highest tiers, threats should target **concepts, not just lives**. Victory isn't ''we survived``---it's *''we changed what it means to be human in this world.``*

\textbf{Concept-threat examples (with GM guidance):}
\begin{itemize}
  \item \textbf{The Memory Plague}: A curse that erases specific memories from anyone who hears its signal (a song, a scent, a symbol). 
        \begin{itemize}
          \item \textit{Player goal}: Find the source and break the curse. 
          \item \textit{If players fail}: The plague spreads---but not randomly. It *targets* memories of kindness, making society crueler and more distrustful. Cities fortify not against invaders, but against *each other*. 
          \item \textit{If players succeed (with cost)**: The plague is halted---but to break it, they had to *forget their own happiest memory*. Describe: ''The sigil crumbles---but as the last echo fades, you realize you can't recall your first kiss. The grief is sharp\dots but so is the clarity. *`Was it worth it*'?`` 
          \item \textit{GM Tip:} Don't just say ''memories are lost``---show the *ripple effect*. A player who forgot their mentor's voice might now miss subtle cues in negotiations. 
        \end{itemize}
  \item \textbf{The God-Law}: A Patron attempts to rewrite the fundamental laws of magic (e.g., ''All healing now requires a blood sacrifice``). 
        \begin{itemize}
          \item \textit{Player goal}: Stop the ritual or find a loophole in the new law. 
          \item \textit{If players fail}: Magic becomes more potent---but every heal requires a life. Societies split: the rich buy immortality; the poor die from treatable wounds. 
          \item \textit{If players succeed (with cost)**: The God-Law is halted---but to do it, they had to *accept a permanent geas*: ''You must never heal a wound caused by selfishness.`` Describe: ''The sigil shatters---but as the light fades, you feel the weight settle on your soul. You heal the knight's wound\dots but when the bandit bleeds out, your hands stay clean. *`The law is in your bones*'.`` 
          \item \textit{GM Tip:} Frame the geas as a *narrative permission*: Players *can* still heal---but only when it serves a higher ideal. It's not a penalty; it's a new way to engage with the world. 
        \end{itemize}
  \item \textbf{The Fractured Sky}: A tear in reality where physics behaves oddly (gravity reverses in patches, time loops in spots). 
        \begin{itemize}
          \item \textit{Player goal}: Seal the tear or learn to navigate it. 
          \item \textit{If players fail}: The tear grows---reality becomes unpredictable. Gravity fails randomly; people age backward in patches; dreams bleed into waking life. 
          \item \textit{If players succeed (with cost)**: The tear is sealed---but to do it, they had to *become an anchor for reality*. Describe: ''The fissure closes---but as the light returns, you feel the weight of the world pressing on your chest. You take a step---and the ground *doesn't move* beneath you. *`You are the still point*'.`` 
          \item \textit{GM Tip:} Make the cost a *new way to be in the world*---not a penalty. The player isn't ''weaker``; they're *different*, with unique opportunities and burdens. 
        \end{itemize}
\end{itemize}
\textit{Core principle:} At mythic tiers, the question isn't *''Can we win?``*---it's *''What are we willing to become to win?``*

\section{Running Mythic Scenarios (Tiers IV--V): Grounding the Cosmic in Choice}
\label{sec:mythic-scenarios}
\index{mythic scenarios}

At the apex of power, the game's tension shifts from *''Can I succeed?``* to *''What story do I want to tell---and what am I willing to lose to tell it?``* Keep play grounded by making cosmic stakes *personal* and *irreversible*.

\subsection*{Handling Reality-Altering Magic: The Price of Cosmic Scale**
When a player attempts something cosmic (''I wish the volcano had never erupted``), **never say no**. Instead, offer a compromise where the effect happens---but the price is *felt*, not abstracted.

\textbf{Framework for cosmic compromises:}
\begin{itemize}
  \item \textbf{Sever a connection}: ''You can do it---but it will cost you your bond to this place. You will *never return* here as long as you live.`` 
        \textit{Example}: Stopping the volcano means the character can never step foot in their homeland again---the land itself rejects them. 
  \item \textbf{Instability aura**: ''It works---but for the remainder of the session, the world is unstable. Every roll generates an extra SB.`` 
        \textit{Example}: The volcano never erupts---but now, minor magic risks wild surges (a simple light spell might explode). 
  \item \textbf{Patron's demand**: ''You succeed---but your Patron demands a permanent sacrifice. One of your Talents is erased, or you must take a geas.`` 
        \textit{Example**: The volcano is stopped---but the Oath of Flame \& Light now requires the character to *never draw a weapon in anger*. 
  \item \textbf{Truth fragmentation**: ''It works---but your perception of reality is now fractured. Once per session, the GM may describe a scene that *isn't really there*.`` 
        \textit{Example**: The volcano is halted---but now, the character sees ghosts of those who died in the eruption. Are they real? Or is their mind breaking? 
\end{itemize}
\textit{GM Tip:} Always tie the price to the player's *identity*, not just their sheet. Losing a homeland isn't just ''no more safehouse``---it's ''the place that knew your name is gone.`` Losing a Talent isn't just ''--2 dice``---it's ''the thing that made you *you* is silent.``

\subsection*{Player Agency at Cosmic Scale: Let the World Break**
\begin{tcolorbox}[
  colback=green!5!white,
  colframe=green!75!black,
  title=At Tier V, Every Choice May Erase an Alternative Timeline,
  fonttitle=\bfseries
,breakable]
\index{mythic scenarios!sacrifice}
Embrace this. Let players narrate what they are saving (or sacrificing) when they act.

\medskip
\noindent\textit{Example:} ''I stop the ritual---but I let the rival escape.`` That rival becomes a future problem. ''I save the city---but I lose my connection to my Patron.`` Now the player has a new arc: how to regain favor or find a new source of power.

\medskip
\noindent\textbf{Let the world break.} If the climax involves a collapsing reality, the battlefield itself can be a timer. Example: \emph{Reality Fracture [8]}. Each time a major event occurs (a spell, a death, an oath), tick the timer. When it fills, the scene shifts---time loops, gravity reverses, or the characters meet themselves from another timeline.
\end{tcolorbox}

\textbf{Reality Fracture [8] in action:}
\begin{itemize}
  \item \textbf{Tick when}: 
        \begin{itemize}
          \item A character invokes a Patron's name in desperation (''*`Oath of Flame \& Light---help me*'!``) 
          \item A major magical effect is attempted (a spell that would normally fail) 
          \item A character makes a life-or-death choice (e.g., ''I jump in front of the arrow to save them``) 
        \end{itemize}
  \item \textbf{At 4 segments}: Strange echoes appear---describe them sensorily: 
        \begin{itemize}
          \item Ghostly duplicates flicker at the edge of vision (''You swear you saw yourself turn the corner---*but you didn't*``) 
          \item Time slips cause déjà vu so strong it hurts (''You *know* you've had this conversation before---*but you haven't*``) 
          \item Local gravity hiccups make objects float for a breath before falling 
        \end{itemize}
  \item \textbf{At 8 segments}: The scene transforms---describe the shift *through player action*: 
        \begin{itemize}
          \item \textit{If a player tries to run}: ''Your feet push off---but instead of moving forward, you *slide sideways* as if the floor is ice. Gravity has *shifted*.`` 
          \item \textit{If a player tries to speak}: ''Your words come out backward---and as you hear them, you realize you're *apologizing for something you haven't done yet*.`` 
          \item \textit{If a player tries to heal}: ''As your hands glow, the wound *closes*---but the skin *around* it blackens and crumbles. Healing here *demands* a trade.`` 
          \item \textit{If a player stands still}: ''The air hums---and for a heartbeat, you *see the battle as it happened ten minutes ago*. Then the vision fades, leaving you shaken.`` 
        \end{itemize}
  \item \textit{GM Tip:} Never just say ''gravity reversed.`` Describe *what the player experiences* when they try to act. Make them *feel* the world's rules changing. 
\end{itemize}
\textit{This turns a mechanic into a visceral, story-defining moment.}

\subsection*{Let Players Name the Cost**
When a cosmic choice arises, **ask the player to state the cost *in their own words***---then hold them to it. This makes the sacrifice feel earned, not imposed.

\textbf{Example:} 
\begin{itemize}
  \item Player declares: ''I throw myself on the grenade to save the squad.`` 
  \item You ask: ''What does this cost you *beyond* your life?`` 
  \item Player responds: ''I'll never see my sister grow up. I won't be there to teach her how to fight.`` 
  \item You narrate the outcome: ''The blast throws you back---but as the smoke clears, you see your squad safe. *And* you feel the absence where your sister's laughter should be in your dreams. *You keep your promise*.`` 
\end{itemize}
\textit{This turns a tactical choice into a defining character moment---where the player *chooses* what they sacrifice.}

\section{Managing Cross-Patron and Invoker Conflicts: When Gods Disagree}
\label{sec:cross-patron-conflicts}
\index{Patrons!cross-Patron conflicts}\index{Invoker!Symbol friction}

When a character serves multiple Patrons (e.g., an Invoker with several Symbols, or a Runekeeper who also carries a Symbol), their obligations can create **narrative tension**---not just mechanical bookkeeping. Your job is to make this tension *felt*, not just tracked.

\subsection*{Signs of Cross-Resonance: Make the Conflict Visible**
Don't just track Obligation segments---**describe the friction** as it builds. Players should *feel* the pull of competing loyalties.

\textbf{Observable signs of growing conflict (describe these as they happen):}
\begin{itemize}
  \item \textbf{Physical sensations}: ''Your Symbols grow warm when near each other---*`Like they're arguing*'.`` or ''A headache builds behind your eyes when you use both Rites.`` 
  \item \textbf{Dreams and omens}: ''Your dreams show two figures---one in light, one in shadow---*`pointing at you*'.`` or ''The omens conflict: one Patron whispers *`flee*'; the other screams *`stand*'.`` 
  \item \textbf{Rite costs and effects}: ''When you invoke Patron A's Rite, it costs +1 Obligation *because Patron B is watching*.`` or ''Patron B's Rite works---but the light it casts feels *thin and doubtful*.`` 
  \item \textbf{Roleplay urges**: ''You feel compelled to *defend* Patron A's followers---even when they're wrong.`` or ''You catch yourself *apologizing* to Patron B's symbols after using Patron A's power.`` 
\end{itemize}
\textit{GM Tip:} When describing these, use sensory language. Don't say ''Obligation conflict``---say ''The oath-scar on your palm burns when you touch the other Symbol.``

\subsection*{Resolving Conflicts: Make the Choice Matter**
When a character invokes two conflicting Patrons in the same scene, **avoid mechanical penalties**. Instead, make the resolution a *story pivot* where the player must pick a side---or try to bridge the divide.

\textbf{Frameworks for resolution (GM-guided, not GM-imposed):}
\begin{itemize}
  \item \textbf{Devil's Bargain (Player Choice)**: 
        \begin{itemize}
          \item Offer: ''You can use both Rites---but mark 2 Fatigue, and I gain 2 SB to spend on showing why this duality is dangerous.`` 
          \item \textit{Player choice**: Accept the bargain (gain power, accept consequence) or decline (lose the power, avoid the immediate cost). 
          \item \textit{GM Tip:} Make the 2 SB *meaningful*: Spend them to show the Patrons' displeasure manifesting (''As you lower your hands, the altar cracks---*`You serve two masters*'.`` and ''A raven croaks three times---*`The Watcher is displeased*'.``) 
        \end{itemize}
  \item \textbf{Bridging the Divide (Player Ingenuity)**: 
        \begin{itemize}
          \item Offer: ''You can try to serve both---but it will take effort. Describe *how* you reconcile them, and I'll set a DV based on your creativity.`` 
          \item \textit{Player action**: They might fast and meditate, offer a joint sacrifice, or find a third path that honors both. 
          \item \textit{GM adjudicates**: Success means they avoid immediate conflict (no Obligation tick, no Fatigue); failure means they offend one Patron (Obligation +1, etc.). 
          \item \textit{Why it works**: It's not ''I can do both``---it's *''I can try to serve both masters---but it requires me to *become* something new*.`` 
        \end{itemize}
  \item \textbf{Forced Choice (Narrative Inevitability)**: 
        \begin{itemize}
          \item Offer: ''You cannot serve both---not without breaking. Describe *which Patron you choose to follow fully*, and what you sacrifice to do so.`` 
          \item \textit{Player action**: They pick a Patron, narrate why, and describe what they give up (e.g., ''I follow the Oath---but I throw the Symbol of the Traveler into the sea. *`My restless heart is still*'.``) 
          \item \textit{GM narrates the consequence**: The chosen Patron's power flows freely; the neglected Patron's Symbol grows cold and heavy, or their omens turn hostile. 
          \item \textit{GM Tip:} Make the sacrifice *hurt*---not just mechanically, but narratively (''As the Symbol leaves your hand, you feel the loss like an amputated limb. *`I chose the oath\dots but I miss the road*'.``) 
        \end{itemize}
\end{itemize}
\textit{Core principle:} Cross-Patron conflict shouldn't feel like a resource drain---it should feel like *a character-defining moment where loyalties are tested*.

\subsection*{Narrative Resolution: Let the Tension Drive Arcs**
Encourage players to roleplay the tension---not just resolve it mechanically. Use these moments to seed long-term growth.

\textbf{How to turn tension into arcs:}
\begin{itemize}
  \item \textbf{The ''Almost Bridged`` arc**: A player keeps trying to serve both Patrons but keeps failing. Each failure teaches them something new about themselves or the Patrons. 
        \textit{Example:} A Runekeeper of the Oath \& The Traveler keeps trying to balance oath-keeping with wayfinding---but each time, they either break an oath or get hopelessly lost. Over time, they realize: *`My oath isn't about rules---it's about who I protect*'. Their evolution isn't mechanical; it's *philosophical*. 
  \item \textbf{The ''Chosen Path`` arc**: A player picks a Patron---and now must live with the consequences of neglecting the other. 
        \textit{Example:} An Invoker chooses the Traveler's Symbol---and now the Oath of Flame \& Light sends omens: dreams of fire, urges to seek conflict. Their arc isn't about gaining power---it's about *learning to live with the god they turned away from*. 
  \item \textbf{The ''Third Way`` arc**: A player discovers a synthesis that honors both Patrons without full allegiance to either. 
        \textit{Example:} A Cantor finds that their songs gain power when they sing *about* the tension between two Patrons---not choosing one. Their arc isn't about power---it's about *becoming a translator between worlds*. 
\end{itemize}
\textit{GM Tip:} In your session notes, track these ideological shifts---not just XP gains. A character who redefines their relationship with a Patron has changed more than one who gained +1 die.

\section{Player-Managed Modules at High Tiers: Trust the Player, Verify the Fiction}
\label{sec:player-managed-modules-high-tier}
\index{player-managed modules!high tier}\index{GM!player-managed}

At high tiers, players track more resources (multiple Obligation segments, Corruption timers, Asset conditions). **As GM, your role isn't to audit---it's to spot-check and enforce narrative consequences when the fiction demands it.** Trust the player's tracking---*but verify that the fiction matches the numbers*.

\subsection*{Player Responsibilities (Reinforce These)**
Remind players of their duties---not as rules, but as *how they keep the story honest*:
\begin{itemize}
  \item \textbf{Declare thresholds out loud**: ''My Obligation to the Traveler just hit capacity---I can't invoke another Rite without clearing it.`` (Lets you prepare consequences) 
  \item \textbf{Maintain Asset states visibly**: If your Safehouse is Neglected, note it on your sheet and say: ''I'll need to tend it before I can use it.`` (Prevents ''I forgot`` excuses) 
  \item \textbf{Track Corruption Blooms actively**: When your Corruption Timer fills, *choose or roll* the new trait---and be ready to roleplay its cost immediately. (Prevents ''I forgot I have this trait`` breaks) 
\end{itemize}
\textit{Player Tip:} Your tracking isn't for the GM---it's for *you* to know when your character is at a breaking point.

\subsection*{GM Spot-Checking: Verify the Fiction, Not the Numbers**
Don't hunt for math errors---instead, **ask questions that force players to connect their numbers to the fiction**. If the fiction matches, trust the math.

\textbf{Effective spot-check questions (ask casually, not confrontationally):}
\begin{itemize}
  \item \textbf{For Obligation/Corruption}: ''You say your Oath of Flame \& Light Obligation is at 5/6---what's the *last thing* you did that made the Oath uneasy?`` 
        \textit{If they describe a meaningful action (e.g., ''I lied to protect a secret``), trust the number. If they say ''I don't know`` or give a vague answer, the number may be drifting from fiction---ask them to reconcile it. 
  \item \textbf{For Asset states}: ''Your Safehouse is Neglected---what's *one thing* you've neglected this week? The chains? The stores? The landlord's trust?`` 
        \textit{If they give a specific, fiction-based answer (e.g., ''I haven't paid the landlord in two weeks``), trust the state. If they say ''I don't know`` or ''it's just old,`` the state may be drifting. 
  \item \textbf{For Corruption Blooms**: ''Your new trait is *`Whispers in the blood*'---how did it show up *today*?`` 
        \textit{If they describe a moment (e.g., ''I almost lied to the guard---but the whisper stopped me``), trust the bloom. If they say ''I haven't noticed it,`` the bloom may be inactive or misapplied. 
\end{itemize}
\textit{GM Tip:} If the fiction and numbers align, *never* correct the player's tracking---even if you suspect a minor error. The story matters more than the sheet. If they drift, *gently* guide them back to fiction---never punish a slip.

\subsection*{Enforcing Consequences: When Fiction and Numbers Diverge**
If a player's declared action *contradicts* their tracked resources (e.g., they claim to use a Neglected Asset), **enforce the fiction---not the number**. 

\textbf{How to handle divergence:}
\begin{itemize}
  \item \textbf{Player says**: ''I use my Safehouse to start Dominant.`` 
  \item \textbf{You know**: They marked it Neglected last session (and haven't tended it). 
  \item \textbf{Do not say**: ''Your sheet says Neglected---you can't.`` 
  \item \textbf{Do say**: ''As you reach for the lever, you notice the chains are rusted stiff---and the landlord's face is grim when you meet their eyes. *`You've been neglecting us*'.`` Then spend 1 SB to show the consequence (''The lever sticks---but you muscle it open. For one beat, the pursuit is negated\dots but the landlord follows you, demanding rent.``) 
  \item \textit{Result**: The player *feels* the neglect---they don't just lose a mechanic; they experience the narrative cost. 
\end{itemize}
\textit{GM Tip:} Never let a player ''win`` by ignoring their tracked state. Make them *earn* the benefit through fiction---or pay the price for ignoring it.

\section{Complex Scenarios: Weaving Systems Together}
\label{sec:complex-scenarios}
\index{complex scenarios}

At high tiers, the best scenarios don't isolate mechanics---they weave them together so players must manage *multiple systems* to succeed. Design scenarios where success requires juggling fiction, resources, timers, and relationships.

\subsection*{Heists: Player-Designed Plans with Systemic Tension**
Instead of planning the heist for them, **ask players to define how their character contributes**---then build tension around *how their role strains their resources and relationships*.

\textbf{GM Procedure:}
\begin{enumerate}
  \item \textbf{Ask each player}: ''How does your character contribute to the heist? Be specific---what skill, relationship, or resource do you bring?`` 
        \textit{Example answers*: 
                \begin{itemize}
                  \item *The Face*: ''I use my Persuasion and Sway to talk our way past the outer guards.`` 
                  \item *The Ghost*: ''I use my Stealth and a Favored Spirit to slip through the wards unseen.`` 
                  \item *The Muscle*: ''I use my Melee and a War Follower to handle any resistance inside.`` 
                  \item *The Brain*: ''I use my Lore and Investigation to find the weak point in the vault's seal.`` 
                \end{itemize}
  \item \textbf{Build a Heist Progress [8]} timer**: Each player's action advances the timer---but each action also strains their resources or relationships. 
        \begin{itemize}
          \item *The Face's Persuasion roll*: Success advances the timer---but if they fail, they mark 1 Obligation to the noble they lied to (they promised a favor for access). 
          \item *The Ghost's Stealth roll*: Success advances the timer---but if they use a Favored Spirit, they advance their Corruption Timer (spirit magic risks Patron Ire). 
          \item *The Muscle's Melee roll*: Success advances the timer---but if they take Harm, they must clear Fatigue first (the roller-coaster effect). 
          \item *The Brain's Lore roll*: Success advances the timer---but if they fail, they gain 1 Boon (a Partial means they learned something useful). 
        \end{itemize}
  \item \textbf{Spend SB from misses to introduce security system obstacles**: 
        \begin{itemize}
          \item A Miss on The Face's roll: Spend 1 SB to show the guard *wasn't* fooled (''His eyes narrow---*`You're lying. What's the real reason you're here*'?`` Then advance \emph{Guard Suspicion [4]} timer +1). 
          \item A Miss on The Ghost's roll: Spend 2 SB to show the ward reacted (''The sigils flare---as you dodge, you feel the ward's tug on your spirit. *`You're not welcome here*'.`` Then advance \emph{Ward Backlash [4]} timer +1 and mark 1 Fatigue from spiritual strain). 
          \item A Miss on The Muscle's roll: Spend 1 SB to show resistance was expected (''The door bursts open---not with thieves, but with the *captain's personal guard*.`` Then advance \emph{Internal Threat [6]} timer +1). 
          \item A Miss on The Brain's roll: Spend 1 SB to show the seal was tougher than expected (''The etchings are deeper than the schematics showed---but you see a maker's mark: *`This was forged by the Black Ledger*'.`` Then note that they now have a Debt Timer to the Black Ledger). 
        \end{itemize}
  \item \textbf{At the climax, make success require a painful choice**: 
        \textit{Example**: The vault is open---but to take the artifact, they must trigger a trap that will seal them in unless someone stays behind. 
                \begin{itemize}
                  \item *If they sacrifice a player*: Describe the character's final act (''*The Ghost* steps forward---*`Go. I'll hold the ward*'.`` Then resolve their fate via the Roller-Coaster effect: they take Harm to clear Fatigue, but the Harm overwhelms them. 
                  \item *If they try to disarm the trap*: It's a High-Stakes Roll (DV 6, Desperate)---success means they all escape; failure means they're trapped and the artifact is lost. 
                \end{itemize}
\end{itemize}
\textit{Why this works}: The heist isn't just ''roll to succeed``---it's ''how do your choices *change you* to get here?``

\subsection*{Mass Combat: Abstract the Horror, Keep the Humanity**
Instead of rolling for every soldier, **treat armies as single units with visible timers and personalities**---but keep room for personal heroism.

\textbf{GM Procedure:}
\begin{enumerate}
  \item \textbf{Define each army as a unit with**: 
        \begin{itemize}
          \item \textbf{A name and purpose} (*''The Iron League's 3rd Legion: disciplined, but low on supplies``*) 
          \item \textbf{Visible timers} reflecting their state (*\emph{Morale [6]}*, \emph{Supply [4]}*, \emph{Discipline [4]}*) 
          \item \textbf{A signature tactic} (*''They advance in shield walls, grinding forward like a glacier``*) 
          \item \textbf{A breaking point} (*''When \emph{Morale} hits 0, they rout; when \emph{Supply} hits 0, they begin to starve``*) 
        \end{itemize}
  \item \textbf{Player actions affect army timers through fiction---not just math**: 
        \begin{itemize}
          \item \textbf{Commanding} (Presence + Command): Success advances \emph{Morale} +1 or retreats \emph{Enemy Morale} +1 (based on fiction: an inspiring speech vs. a terrifying threat) 
          \item \textbf{Sabotaging} (Wits + Craft): Success retreats \emph{Enemy Supply} +1 or advances \emph{Own Supply} +1 (based on fiction: poisoning wells vs. repairing wagons) 
          \item \textbf{Rallying the troops} (Presence + Sway): Success advances \emph{Morale} +2 (based on fiction: a shared memory, a promise kept) 
          \item \textbf{Personal combat} (Attack rolls): Happens *within* the mass conflict---but only if the player chooses to engage directly. 
        \end{itemize}
  \item \textbf{When an army timer fills, describe the breaking point *through fiction*---not just a number**: 
        \begin{itemize}
          \item \emph{Morale [6]} fills: ''As the horns sound, you see it---their shoulders slump. Not in fear, but in *exhaustion*. They've marched too long, fought too hard. The advance stalls---not broken, but *done*.`` 
          \item \emph{Supply [4]} fills: ''The wagons creak empty---but worse, you see the men chewing on leather straps. Starvation has *begun*.`` 
          \item \emph{Discipline [4]} fills: ''The line wavers---not from fear, but from *doubt*. Murmurs spread: *`Why are we fighting for this*'.`` 
        \end{itemize}
  \item \textbf{Allow personal heroism to create micro-timers within the mass conflict**: 
        \begin{itemize}
          \item A player who holds a breach against overwhelming odds starts a \emph{Hold the Line [4]} timer. 
          \item When it fills: ''They've held for ten minutes---but as the enemy presses, you see their spear arm tremble. *`I can't…*'.`` (They fall---but their action gave the army time to retreat) 
        \end{itemize}
\end{itemize}
\textit{Why this works}: Players don't just ''win the battle``---they *feel* the weight of command, the horror of attrition, and the glory of standing firm when others break.

\subsection*{Political Intrigue: Reputation as a Living Resource**
Track a \emph{Reputation Timer [6]} for the party in a given court---not as a score, but as a *narrative temperature*.

\textbf{GM Procedure:}
\begin{itemize}
  \item \textbf{Define what moves the timer**: 
        \begin{itemize}
          \item \textbf{Advances (+1)}: Public deeds that build trust (e.g., ''You return the stolen artifact to the temple``), favors granted without expectation of return (''You heal the sick child---and ask nothing in return``), or speeches that resonate (''Your words about justice make the crowd nod slowly``). 
          \item \textbf{Retreats (-1)}: Actions that erode trust (e.g., ''You're seen taking a bribe from the merchant``), broken promises (''You promised to meet the duke---but you were at the tavern``), or divisive rhetoric (''Your speech calls the duke a tyrant---*`and his supporters are fools*'``). 
        \end{itemize}
  \item \textbf{When the timer advances, describe the shift *in the crowd's behavior***: 
        \begin{itemize}
          \item \textbf{+1 segment}: ''As you speak, you notice heads turning---not with suspicion, but with *interest*. A merchant murmurs to his friend: *`There's sense in what they say*'.`` 
          \item \textbf{+2 segments**: ''The bailiff waves you through without checking your papers---*`You've earned trust here*'.`` 
          \item \textbf{+3 segments**: ''As you enter the hall, the crowd parts---not to let you pass, but to *honor* you. An elder steps forward: *`We've waited for you*'.`` 
        \end{itemize}
  \item \textbf{When the timer retreats, describe the shift *in how others treat them***: 
        \begin{itemize}
          \item \textbf{-1 segment**: ''As you approach the gate, the guard's hand drifts to his sword hilt---*`We'll see*'.`` 
          \item \textbf{-2 segments**: ''The innkeeper sees you come in---and his smile vanishes. *`Not tonight*'.`` 
          \item \textbf{-3 segments**: ''As you pass, you hear a mother pull her child closer---*`Stay away from them*'.`` 
        \end{itemize}
  \item \textbf{When the timer fills (6/6), describe the ''trusted`` state *as an ongoing reality*---not just a one-time bonus**: 
        \begin{itemize}
          \item \textbf{Benefits not granted as mechanics**: Instead of ''+2 die,`` describe: ''The magistrate looks at your case---and instead of asking for proof, he says: *`I believe you*'.`` or ''The captain falls in step beside you---*`I've got your back*'.`` 
          \item \textbf{Benefits not permanent**: Make it clear this trust is *fragile*---it can retreat if they act against it. 
          \item \textbf{Why it works**: Reputation isn't a bonus---it's *how the world sees them*. When trusted, they don't need to *prove* themselves; they're *given* the benefit of the doubt. 
        \end{itemize}
  \item \textbf{When the timer empties (0/6), describe the ''distrusted`` state *as an ongoing reality***: 
        \begin{itemize}
          \item \textbf{Not just ''---2 die``**: Describe: ''The magistrate glances at your case---and instead of hearing you out, he says: *`We've heard this before*'.`` or ''Soldiers avoid your patrol route---not fear, but *disgust*.`` 
          \item \textbf{Why it works**: When distrusted, they don't just get penalties---they're *actively worked against*. 
        \end{itemize}
\end{itemize}
\textit{Why this works}: Players don't just ''manage a timer``---they *live* in a world where their reputation *changes how they're treated*, moment by moment.

\section{When to Say Yes, No, and Yes, But: Advanced Rulings}
\label{sec:advanced-rulings}
\index{GM rulings!advanced}

At high tiers, players will attempt things that strain normal bounds. **Your job isn't to say ''yes`` or ''no``---it's to say ''yes, and here's what it costs`` or ''no, but here's how you can get close.``** Never shut down creativity---channel it into meaningful tension.

\subsection*{Compromise Solutions: The ''Yes, And`` Framework**
Instead of a flat ''yes`` or ''no,`` offer a compromise where the effect happens---but the price is *narrative*, not mechanical. 

\textbf{Framework for compromise (choose based on fiction):}
\begin{itemize}
  \item \textbf{Yes, and it alters your relationship**: 
        \begin{itemize}
          \item \textit{Example**: ''You can stop the ritual---but to do it, you must *publicly renounce* your Patron. Describe how you do it, and what their reaction is.`` 
          \item \textit{GM Tip:} Make the alteration *specific and painful*---not just ''you lose favor.`` (''As you speak the words of renunciation, you feel the bond snap like a dry twig. *`I am no longer yours*'.``) 
        \end{itemize}
  \item \textbf{Yes, and it demands a sacrifice**: 
        \begin{itemize}
          \item \textit{Example**: ''You can win the battle---but to do it, you must *leave someone behind*. Describe who stays, and why.`` 
          \item \textit{GM Tip:} Make the sacrifice *irreversible and personal*---not just ''you lose a Follower.`` (''As you ride away, you see *the Ghost* kneeling in the mud, shielding the wounded with their body. *`Go---tell them I died holding the line*'.``) 
        \end{itemize}
  \item \textbf{Yes, but it takes longer---and here's the cost**: 
        \begin{itemize}
          \item \textit{Example**: ''You can forge the weapon---but it will take a month of nonstop work. Describe what you *neglect* during that time (family? duties?).`` 
          \item \textit{GM Tip:} Make the neglect *visible and costly*---not just ''you're busy.`` (''As the hammer falls, you realize you haven't visited your sister's grave in three weeks. *`She'd be ashamed*'.``) 
        \end{itemize}
  \item \textbf{Yes, but the GM gains SB to spend on showing why it's dangerous**: 
        \begin{itemize}
          \item \textit{Example**: ''You can attempt the spell---but I gain 3 SB to spend on showing why the world resists you.`` 
          \item \textit{GM Tip:} Spend the SB to show *specific, escalating danger* (''As you begin the chant, the air smells of ozone---*`This is drawing power from somewhere wrong*'.`` Then spend SB to show the power flickering, then backfiring, then attracting a hostile spirit). 
        \end{itemize}
\end{itemize}
\textit{GM Tip:} Always frame the compromise as: *''You can do X---but here's what it *costs you in the story*.``* Never let it feel like a mechanical trade-off.

\subsection*{Saying ''No`` Without Shutting Down: The ''No, But`` Framework**
If a request is truly impossible (e.g., ''I kill the god with a single arrow``), **don't just say no---offer a path forward that honors the intent**.

\textbf{Framework for ''no, but``:}
\begin{enumerate}
  \item \textbf{Explain why it's impossible *in the fiction***: 
        \begin{itemize}
          \item \textit{Not**: ''The rules say you can't.`` 
          \item \textit{Do**: ''The god isn't a physical being---it's the *idea* of justice itself. To harm it, you'd have to unmake every act of justice ever done in this world.`` 
        \end{itemize}
  \item \textbf{Offer a path that gets *close* to the intent**: 
        \begin{itemize}
          \item \textit{Example**: ''But if you find the *Heartstone*---the crystal where the first just act was sealed---and *shatter it*, you could weaken the god's influence. You'd need to gather three relics that *represent unjust acts*.`` 
          \item \textit{GM Tip:} Make the path *demanding but possible*---not a easy way out, but a meaningful quest that tests their commitment to the goal. 
        \end{itemize}
  \item \textbf{Let the dice decide how close they get---but hedge the outcome**: 
        \begin{itemize}
          \item \textit{Example**: ''Set DV 8 for Wits + Lore (Desperate Position). If they succeed, they find the Heartstone---but learn that shattering it would *unmake all justice, even the good*. If they fail, they get a clue to where the Heartstone *might* be.`` 
          \item \textit{GM Tip:} On success, don't let them ''win``---give them a *Pyrrhic victory* where success creates a new problem. On failure, give them *meaningful progress* (a clue, an alliance, a lesson) that keeps the story moving. 
        \end{itemize}
\end{itemize}
\textit{GM Tip:} Never let a ''no`` feel like a dead end. Always turn it into: *''You can't do X---but here's how you can get closer, and what you'll learn along the way.``*

\subsection*{Let the Dice Decide---but Always Add Narrative Weight**
For requests that are *possible* but risky (e.g., ''I try to bluff my way past the dragon``), **use an extreme DV and Desperate Position**---but *always* add narrative weight to the outcome.

\textbf{Framework for dice-checked risky actions:}
\begin{itemize}
  \item \textbf{Set DV 6+ and Desperate Position**: 
        \begin{itemize}
          \item \textit{Example**: ''Bluffing a dragon? That's DV 8, Desperate.`` 
          \item \textit{GM Tip:} Make the Desperate re-roll *mean something*---not just a mechanic. (''As you speak, you feel the dragon's gaze like a physical weight. *`One wrong word, and you're ash*'.``) 
        \end{itemize}
  \item \textbf{On success, describe how they succeeded *through fiction*---not just a number**: 
        \begin{itemize}
          \item \textit{Example**: ''You don't bluff the dragon---you *tell it a truth* it hasn't heard in centuries: *`Your hoard is beautiful, but it's lonely*'.`` The dragon hesitates---then lets you pass. 
          \item \textit{GM Tip:} Make the success *about the player's character*---not just beating the roll. (''As you speak, you see your own loneliness reflected in the dragon's eyes. *`We're both lonely*'.``) 
        \end{itemize}
  \item \textbf{On failure, describe how they failed *through fiction*---not just a number**: 
        \begin{itemize}
          \item \textit{Example**: ''The dragon doesn't roar---it *laughs*, a sound like breaking stone. *`You think words can stop me?*'.`` Then it breathes fire---not to kill, but to *scorch them black as a warning*. 
          \item \textit{GM Tip:} Make the failure *a lesson*---not just a penalty. (''As the smoke clears, you see the dragon's amusement in its eyes. *`Go---and tell others what happens to liars*'.``) 
        \end{itemize}
\end{itemize}
\textit{GM Tip:} Never let the dice be the *whole* story. The roll decides *if* they succeed---but the *fiction* decides *what it means*.

\section{Quick Reference: Advanced GM Toolkit}
\label{sec:gm-advanced-quick-ref}
\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Advanced GM Techniques Quick Reference,breakable]
\begin{itemize}
  \item \textbf{SB Management}: 
        \begin{itemize}
          \item Spend SB to advance *multiple* timers/fronts (not one big complication) 
          \item Use banked SB *only* for complications tied to a character's background/Patron 
          \item Never hoard SB---if the fiction needs a complication, spend it 
        \end{itemize}
  \item \textbf{Deck of Consequences}: 
        \begin{itemize}
          \item Synthesize draws into *one narrative twist* (not a list of effects) 
          \item Use the deck for downtime worldbuilding, flashback seeds, and montage triggers 
          \item Customize suit meanings to your campaign's themes 
        \end{itemize}
  \item \textbf{Thresholds}: 
        \begin{itemize}
          \item Respond to threshold crossings with *narrative prompts* (not mechanics tables) 
          \item Use Threshold Instability timers for liminal locations---tick when players cross without narrative respect 
          \item Let player-specified crossing methods set Position/Effect 
        \end{itemize}
  \item \textbf{Faction Grand Strategy}: 
        \begin{itemize}
          \item Represent factions with 1--3 *narrative* timers (e.g., \emph{Paymaster's Patience}, not \emph{Timer 3}) 
          \item Adjudicate player actions through the faction's *values and current state* (not just math) 
          \item Run faction turns between sessions: player orders → GM adjudication → ''News from the Front`` 
          \item Make neglect escalate in *faction-specific ways* (not generic penalties) 
        \end{itemize}
  \item \textbf{Advanced Timers}: 
        \begin{itemize}
          \item Design interacting timers (see-saw, threshold gating) and cascading timers (threshold breach, escalating threat) 
          \item Hide consequences behind success *only* when the fiction makes it inevitable 
          \item When a timer fills, describe the shift *through player senses and actions* 
        \end{itemize}
  \item \textbf{Mythic Scenarios}: 
        \begin{itemize}
          \item For cosmic attempts, offer compromises where the effect happens---but the price is felt (lost connection, instability aura, Patron's demand) 
          \item Let the battlefield itself be a timer (e.g., \emph{Reality Fracture [8]}) that shifts reality when full 
          \item Ask players to state the cost of their choice *in their own words*---then hold them to it 
        \end{itemize}
  \item \textbf{Cross-Patron Conflicts}: 
        \begin{itemize}
          \item Describe friction through physical sensations, dreams, omens, and Rite changes 
          \item Resolve via Devil's Bargain (player choice), Bridging the Divide (player ingenuity), or Forced Choice (narrative inevitability) 
          \item Use tension to seed long-term arcs (almost bridged, chosen path, third way) 
        \end{itemize}
  \item \textbf{Player-Managed Modules}: 
        \begin{itemize}
          \item Trust player tracking---but spot-check by asking how the fiction matches their numbers 
          \item Never correct tracking if fiction aligns; if mismatched, enforce the *fiction* (not the number) 
          \item Make them declare thresholds, maintain Asset states visibly, and track Corruption Blooms actively 
        \end{itemize}
  \item \textbf{Compromise Solutions}: 
        \begin{itemize}
          \item Yes, and it alters your relationship 
          \item Yes, and it demands a sacrifice 
          \item Yes, but it takes longer---and here's the cost 
          \item Yes, but the GM gains SB to spend on showing why it's dangerous 
        \end{itemize}
  \item \textbf{When Saying ''No``}: 
        \begin{itemize}
          \item Explain why it's impossible *in the fiction* 
          \item Offer a path that gets close to the intent 
          \item Let the dice decide---but hedge the outcome (success = Pyrrhic victory; failure = meaningful progress) 
        \end{itemize}
  \item \textbf{Core Principle}: 
        \begin{itemize}
          \item \textbf{Every mechanical interaction must serve the story**---never let the math override the fiction 
          \item \textbf{At high tiers, complexity > power**: The challenge isn't in stronger bonuses---it's in navigating webs of loyalty, obligation, and consequence 
        \end{itemize}
\end{itemize}
\end{tcolorbox}

\section{Final Thought}
\label{sec:final-thought-advanced}
\textit{Fate's Edge}'s power at high tiers isn't in bigger numbers---it's in **thicker stories**. The mechanics aren't a ceiling to build against; they're scaffolding to build *upon*. When you spend an SB, don't just add a complication---ask: *''How does this make the player's choice matter more?``* When a timer fills, don't just advance a number---ask: *''How does this change what the player *feels* when they act?``* 

The best advanced GMing doesn't make the world harder to succeed in---it makes success *more meaningful*, failure *more instructive*, and every choice *more fraught with consequence*. Because in the end, it's not about whether they win or lose---it's about **what story they get to tell about what they were willing to owe**.

Now go make the world remember