\chapter{Running Assets and Followers}
\label{ch:running-assets-followers}
\index{assets!running}\index{followers!running}\index{GM!assets}\index{GM!followers}

In \textit{Fate's Edge}, Assets and Followers are not mere mechanical bonuses—they are narrative levers that shape the world between scenes. As GM, your role is to introduce them as earned rewards, adjudicate their benefits \textbf{only when narratively justified}, and use their neglect as a primary source of compelling complications. They thrive when they create tension between player ambition and worldly consequences—not when they become spreadsheets of passive benefits.

\section{Design Principles for GM Use}
\index{assets!design principles}\index{followers!design principles}\index{GM!design}
When managing Assets and Followers in your campaign, remember:  
\begin{itemize}
  \item \textbf{They shift fiction between scenes}: Assets open doors, quiet rumors, or stock larders \emph{off-screen}. Followers act offscreen to gather information or manage responsibilities—\emph{never} autonomously in scenes (that's player agency).  
  \item \textbf{They create hooks}: Every Asset or Follower should suggest \emph{both} opportunities (if tended) \emph{and} complications (if neglected). Neglect isn't "broken"—it's a story waiting to happen.  
  \item \textbf{Stay narrative-first}: Prefer position, scale, and access over flat modifiers. Instead of "+2 die," grant "start \textbf{Dominant} in negotiations with the guild" or "Your contact in the watch means the patrol arrives 10 minutes late."  
  \item \textbf{Never let them replace player agency}: If an Asset or Follower solves a problem too cleanly, add a cost (obligation, timer advance, or narrative complication). Their value is in \textit{how} they enable player action, not in removing risk.  
  \item \textbf{They scale with tier}: In Tier I, a single safehouse feels significant; in Tier IV, players should manage networks of Assets or high-Cap Followers to match their influence.  
\end{itemize}

\paragraph{What They Are \emph{Not} (GM Guidance)}
Ensure Assets and Followers \emph{never}:  
\begin{itemize}
  \item Act as scene-solvers (e.g., "My safehouse automatically disarms all traps").  
  \item Generate wealth or XP passively (their value is in narrative leverage, not bookkeeping).  
  \item Stack identical effects—if two Followers would grant the same Assist, the GM chooses the \emph{best narrative fit} (not the strongest mechanical bonus).  
\end{itemize}

\section{Assets: GM Procedures for Introduction and Adjudication}
\label{sec:gm-assets-procedures}
\index{assets!procedures}\index{GM!assets}
Assets are properties, charters, networks, or institutions under player control. \textbf{As GM, follow these steps when introducing or adjudicating Assets:}  

\begin{enumerate}
  \item \textbf{Introduce as Narrative Reward}: Never grant Assets via XP alone. Tie them to fiction:  
        \begin{itemize}
          \item "After clearing the bandit camp, the grateful village grants you use of their old mill as a Safehouse."  
          \item "Your successful negotiation earns you a Guild Seat—vote and shop‑rights in the artisan’s quarter."  
        \end{itemize}
        \textit{Never} say "Spend 4 XP for a Safehouse." Instead: "The mill is yours—but it needs repair before it’s useful."  
  \item \textbf{Price for Story Weight, Not Just XP}:  
        \begin{itemize}
          \item In Tier I (0--40 XP), a Minor Asset (4 XP) should feel like a major investment—adjust cost slightly up/down for narrative significance (e.g., a unique heirloom weapon might cost 6 XP as a Minor Asset).  
          \item In Tier IV+ (151+ XP), Minor Assets are pocket change—reserve 4 XP costs for trivial holdings; Standard (8 XP) and Major (12 XP) Assets should feel meaningful.  
        \end{itemize}
  \item \textbf{Always Attach a Hook}: When introducing an Asset, immediately establish:  
        \begin{itemize}
          \item \textbf{What happens if tended?} (e.g., "The informant ring whispers of a shipment tonight.")  
          \item \textbf{What happens if neglected?} (e.g., "If ignored for two sessions, the informants go silent—or worse, start talking to rivals.")  
        \end{itemize}
  \item \textbf{Adjudicate Benefits Only When Fiction Supports It}:  
        \begin{itemize}
          \item \textbf{Free Use}: Once per session, trigger the off-screen benefit \emph{only if} the Asset is maintained and the fiction supports it (e.g., a Safehouse is only useful if prepared and stocked—track via downtime scenes).  
          \item \textbf{Scene Surge}: Spend 1 Boon to turn an Asset into a dramatic on-screen intervention (one beat). \textbf{Limit to one meaningful use per scene}—more than that feels like a deus ex machina. Describe it vividly (see \S\ref{sec:using-assets-followers}).  
          \item \textbf{Never} grant Free Use automatically—ask: "Does the fiction support using this benefit \emph{now}?" If no, require a scene of preparation or a Boon spend.  
        \end{itemize}
  \item \textbf{Use Neglect as Your Complication Engine}:  
        \begin{itemize}
          \item Track Asset condition visibly (Maintained $\rightarrow$ Neglected $\rightarrow$ Compromised).  
          \item When Neglected: Advance a relevant timer (e.g., \emph{Informant Network [4]} ticks +1 per session neglected) or spend 1--2 SB to create complications tied to decay (e.g., "The safehouse landlord demands rent—or eviction.").  
          \item When Compromised: Deny all benefits until fixed via quest or major downtime.  
        \end{itemize}
\end{enumerate}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title={GM Tip: The Asset Neglect Engine},breakable]
\index{assets!neglect}
\textbf{Neglect is your best storytelling tool.} When players ignore an Asset:  
\begin{itemize}
  \item Advance a relevant timer (e.g., \emph{Shop Rent [4]} ticks +1 per session neglected for a Market Front).  
  \item Spend 1--2 SB to create a complication tied to the Asset's decay (e.g., "The Scholar's Cell custodian leaves—research now requires a DV +1 roll").  
  \item Let the fiction degrade the Asset's benefit (e.g., a neglected Healing House now only clears Fatigue 1 once per downtime).  
\end{itemize}
\textit{Never} treat neglect as "nothing happens." A neglected Asset is a ticking time bomb—or a chance for players to re-engage with their world.
\end{tcolorbox}

\section{Followers: GM Procedures for Introduction and Adjudication}
\label{sec:gm-followers-procedures}
\index{followers!procedures}\index{GM!followers}
Followers are named NPCs with a personal relationship to the player character. \textbf{As GM, follow these steps when introducing or adjudicating Followers:}  

\begin{enumerate}
  \item \textbf{Introduce as Narrative Reward}: Never grant Followers via XP alone. Tie them to fiction:  
        \begin{itemize}
          \item "After saving the merchant’s daughter, she swears loyalty—now she’s your Informant."  
          \item "The defeated captain yields—now he follows you as your Mercenary Lieutenant."  
        \end{itemize}
        \textit{Never} say "Spend XP for a Follower." Instead: "They’ll follow you—but only if you prove yourself worthy."  
  \item \textbf{Price for Story Weight, Not Just XP}:  
        \begin{itemize}
          \item A Cap 1 Follower (e.g., a street urchin) should feel accessible early on.  
          \item A Cap 3+ Follower (e.g., a veteran sergeant) should feel like a major investment—adjust XP cost slightly for narrative weight (e.g., a unique spirit ally might cost 3 XP as a Cap 1 Follower due to story significance).  
          \item High-Cap Followers (Cap 4--5) should be rare and hard-won—reserve them for Tier III+ play.  
        \end{itemize}
  \item \textbf{Always Attach a Hook}: When introducing a Follower, immediately establish:  
        \begin{itemize}
          \item \textbf{What happens if tended?} (e.g., "Your Informant whispers of a shipment tonight.")  
          \item \textbf{What happens if neglected?} (e.g., "If ignored for two sessions, your Informant goes silent—or worse, starts talking to rivals.")  
        \end{itemize}
  \item \textbf{Adjudicate Benefits Only Through Player Action}:  
        \begin{itemize}
          \item Followers \emph{do not} act autonomously in scenes. Their benefits come \emph{only} through:  
                \begin{itemize}
                  \item \textbf{Bonds}: Spend 1 Boon to draw on trust/loyalty for insight or support (see \S\ref{sec:bond-activation}).  
                  \item \textbf{Assists}: Spend 1 Boon to grant +1 die to an ally’s roll (max +3 total assist dice per action).  
                  \item \textbf{Offscreen Action}: Handle tasks that would demand PC time (e.g., "Your Informant spends the week gathering rumors—report back next session").  
                \end{itemize}
          \item \textit{Never} say "My Follower scouts ahead"—instead, the player must spend a Boon for Assist or declare a Bond-driven action.  
        \end{itemize}
  \item \textbf{Track Loyalty and Fitness as States}:  
        \begin{itemize}
          \item Instead of numeric timers, note a Follower’s \textbf{Loyalty} (Faithful/Strained/Broken) and \textbf{Fitness} (Ready/Hurt/Down) as one of three states each.  
          \item Change states \emph{only when the fiction demands it} (e.g., after a failed mission, a moral compromise, or taking damage). \textbf{No rolls required}.  
          \item \textbf{GM Guidance}: Use state changes to drive complications:  
                \begin{itemize}
                  \item Loyalty Strained $\rightarrow$ Assist at --1 die; may refuse risky orders.  
                  \item Loyalty Broken $\rightarrow$ Follower leaves or becomes hostile.  
                  \item Fitness Hurt $\rightarrow$ Assist at --1 die; no independent actions.  
                  \item Fitness Down $\rightarrow$ Requires rescue or medical care.  
                \end{itemize}
        \end{itemize}
  \item \textbf{Use Neglect as Your Complication Engine}:  
        \begin{itemize}
          \item If a Follower’s Loyalty becomes Strained or Broken, all delegated Assets immediately become Neglected (see \S\ref{sec:asset-capacity}).  
          \item If a Follower is Compromised, all delegated Assets are lost.  
          \item Neglecting a Follower’s upkeep (XP or downtime scene) degrades their state—\textbf{use this as your complication engine}.  
        \end{itemize}
\end{enumerate}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title={GM Tip: The Follower Neglect Engine},breakable]
\index{followers!neglect}
\textbf{Neglect is your best storytelling tool.} When players ignore a Follower:  
\begin{itemize}
  \item Degrade their Loyalty/Fitness state based on fiction (e.g., "After missing pay for two sessions, your Mercenary Lieutenant’s Loyalty turns Strained—Assist at --1 die").  
  \item If neglected for 2+ sessions, trigger a consequence (e.g., "Your Informant sells your secrets to a rival—spend 1 SB to learn what they revealed").  
  \item Let the fiction reflect the strain (e.g., a Hurt Follower can’t undertake offscreen tasks—they need tending first).  
\end{itemize}
\textit{Never} treat neglect as "nothing happens." A neglected Follower is a relationship in crisis—or a chance for players to rebuild trust.
\end{tcolorbox}

\section{The Interaction Loop: Assets, Followers, and the Core Mechanics}
\label{sec:interaction-loop}
\index{assets!interaction}\index{followers!interaction}\index{core loop}
Assets and Followers integrate seamlessly with Fate’s Edge’s core resolution loop. \textbf{As GM, guide players through this cycle:}  

\begin{enumerate}
  \item \textbf{Player declares intent and approach} (Attribute + Skill).  
  \item \textbf{You state Position and set relevant Timers}.  
  \item \textbf{Player rolls d10s; counts Successes (6+) and Story Beats (1s)}.  
  \item \textbf{Compare to DV $\rightarrow$ use Outcome Matrix}.  
  \item \textbf{Apply consequences and update the fiction}—\textbf{this is where Assets and Followers shine}:  
        \begin{itemize}
          \item \textbf{On Partial or Miss}: Players may spend Boons to activate Assets/Followers (Scene Surge or Bond-driven action).  
          \item \textbf{On Success}: Players may use Assets/Followers offscreen to prepare for next scene (Free Use or downtime action).  
          \item \textbf{Always}: Check Asset/Follower condition—neglect triggers complications \emph{before} applying the roll’s outcome.  
        \end{itemize}
  \item \textbf{Pass the spotlight}—\textbf{but first, remind players of pending upkeep}:  
        \begin{itemize}
          \item "Before we move on—your Safehouse needs tending this downtime. Pay 4 XP or devote a scene to it, or it becomes Neglected."  
          \item "Your Informant’s Loyalty is Strained—they’ll refuse risky orders until you make amends."  
        \end{itemize}
\end{enumerate}

\begin{tcolorbox}[colback=yellow!5!white,colframe=yellow!75!black,title={GM Tip: The Upkeep Reminder},breakable]
\index{assets!upkeep}\index{followers!upkeep}
\textbf{Make upkeep a narrative moment—not bookkeeping.} At scene end:  
\begin{itemize}
  \item Ask: "Who tends their Assets/Followers this downtime? What does that look like in fiction?"  
  \item Let players describe the action (e.g., "I spend the morning greasing the Safehouse chains," "I buy my Follower a drink and apologize for the last mission").  
  \item Only \textit{then* apply mechanical effects: pay XP, change state, or advance neglect timers.  
\end{itemize}
\textit{This turns upkeep from a chore into roleplay—exactly where Assets and Followers belong.}
\end{tcolorbox}

\section{Using Assets and Followers in Play: GM Procedures}
\label{sec:using-assets-followers}
\index{assets!in play}\index{followers!in play}\index{downtime}\index{flashbacks}
\textbf{As GM, your role is to:}  
\begin{itemize}
  \item \textbf{Frame Asset/Follower benefits only when fiction supports it}:  
        \begin{itemize}
          \item A Safehouse is only useful if prepared and stocked (track via downtime scenes).  
          \item A Follower only assists if their Loyalty/Fitness state allows it (track via fiction-driven state changes).  
          \item \textit{Never} grant benefits automatically—ask: "Does the fiction support using this benefit \emph{now}?" If no, require a scene of preparation, a Boon spend (Scene Surge/Bond-driven action), or a narrative cost.  
        \end{itemize}
  \item \textbf{Allow flashbacks only with a modest cost}:  
        \begin{itemize}
          \item To show prior work (e.g., "You spent yesterday preparing the Safehouse"), require 1 Boon \emph{or} 1 downtime scene.  
          \item \textit{Never} allow free exposition—flashbacks must cost something meaningful.  
        \end{itemize}
  \item \textbf{Tie complications directly to Asset/Follower condition}:  
        \begin{itemize}
          \item If Neglected: Spend SB against it first (e.g., "The Safehouse is Neglected—spend 1 SB to advance the \emph{Landlord Dispute [4]} timer").  
          \item If Compromised: Deny benefits until fixed via quest or major downtime.  
          \item If Follower Loyalty Broken: They leave or become hostile—spend 1 SB to show their departure or betrayal.  
        \end{itemize}
  \item \textbf{Spotlight spread}: Rotate which Assets/Followers take the stage so each investment breathes on-screen now and then—\textbf{never} let one Asset or Follower dominate all scenes.  
  \item \textbf{Use them to create obligation and timers}:  
        \begin{itemize}
          \item Leveraging an Asset against faction interests may add Obligation (see \S\ref{ch:world-powers-obligation}).  
          \item Neglecting a Follower may advance a \emph{Trust Eroded [4]} timer.  
          \item Using a Follower for offscreen work may fill segments on a \emph{Reputation [6]} timer.  
        \end{itemize}
\end{itemize}

\paragraph{Example of Play (GM Perspective)}
A player has a \textbf{Safehouse (Minor Asset)} and an \textbf{Informant (Cap 1 Follower)}. \textbf{As GM:}  
\begin{itemize}
  \item \textbf{Scene Start}: Ask: "Is the Safehouse maintained and stocked? Is the Informant tended?"  
        \begin{itemize}
          \item If yes: Grant Free Use to start an entry/exit scene at \textbf{Dominant} position. Describe: "The landlord nods at your signal—'All clear. Proceed to the cellar.'"  
          \item If no (Safehouse Neglected): Describe: "The floorboard creaks differently. Your landlord eyes you over tea: 'Business has been slow. I wonder why.'" Then spend 1 SB to advance \emph{Landlord Dispute [4]} timer +1.  
          \item If Informant Loyalty Strained: Describe: "Your Informant avoids your gaze. 'I’ve got nothing for you this week—business’s been rough.'" They refuse to assist until tended.  
        \end{itemize}
  \item \textbf{During Scene}: When player wants to use Asset/Follower:  
        \begin{itemize}
          \item For Safehouse Scene Surge (1 Boon): Describe: "You grip the hidden lever—'Hold!' The wine rack slides aside, revealing the escape tunnel. For one beat, pursuit is negated—just enough to slip into the alley."  
          \item For Informant Bond-driven action (1 Boon): Describe: "You slide a coin across the table—'Remember the river bend?' Their eyes light up: 'The shipment leaves at midnight. Take the north ridge.'" Grant +1 die on the next Stealth roll.  
        \end{itemize}
  \item \textbf{Scene End}: Remind players of upkeep:  
        \begin{itemize}
          \item "Your Safehouse needs tending—pay 4 XP or devote a scene to avoid Neglect."  
          \item "Your Informant’s Loyalty is Strained—spend downtime mending trust or they’ll go silent next session."  
        \end{itemize}
\end{itemize}

\section{Pacing and Scaling Guidance}
\label{sec:pacing-scaling}
\index{assets!pacing}\index{followers!pacing}\index{GM!pacing}
\textbf{As GM, use Assets and Followers to scale challenge naturally with tier:}  

\begin{itemize}
  \item \textbf{Tier I (0--40 XP)}:  
        \begin{itemize}
          \item Assets are rare and hyper-local (e.g., a single Safehouse, a trusted Informant).  
          \item Followers are low-Cap (Cap 1--2) and deeply personal (e.g., a sibling, a mentor).  
          \item \textit{GM Guidance}: One Asset or Follower should feel significant—neglect creates immediate, personal complications (e.g., "Your Informant won’t help until you repay their favor").  
        \end{itemize}
  \item \textbf{Tier II (41--90 XP)}:  
        \begin{itemize}
          \item Assets begin to show regional reach (e.g., a Guild Seat, a Harbor Office).  
          \item Followers include mid-Cap (Cap 2--3) figures (e.g., a guild factor, a veteran sergeant).  
          \item \textit{GM Guidance}: Assets/Followers should tie into regional politics or trade—neglect invites rivals or invites scrutiny (e.g., "Neglect your Guild Seat for two sessions, and rivals challenge your voting rights").  
        \end{itemize}
  \item \textbf{Tier III (91--150 XP)}:  
        \begin{itemize}
          \item Assets scale to national influence (e.g., a Banking Charter, a Fortress).  
          \item Followers include high-Cap (Cap 3--4) figures (e.g., a bank director, a garrison commander).  
          \item \textit{GM Guidance}: Assets/Followers should tie into national economics or military power—neglect risks invasion, embargo, or revolt (e.g., "Neglect your Fortress for two sessions, and the garrison mutinies").  
        \end{itemize}
  \item \textbf{Tier IV+ (151+ XP)}:  
        \begin{itemize}
          \item Assets are empire-shaping (see Tier III Institutions, \S\ref{sec:tier-iii-assets} in Compendium).  
          \item Followers include Cap 4--5 figures who can manage Assets on your behalf.  
          \item \textit{GM Guidance}: Assets/Followers should tie into imperial succession or continental wars—neglect risks civil war or foreign invasion (e.g., "Neglect your Palace Complex for two sessions, and court factions secede").  
        \end{itemize}
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title={GM Tip: The Scaling Mantra},breakable]
\index{assets!scaling}\index{followers!scaling}
\textbf{At high tiers, Assets and Followers aren’t "more powerful"—they’re "more complex."}  
\begin{itemize}
  \item A Tier V character doesn’t need "a better Safehouse"—they need "a Safehouse network with competing landlords and shifting loyalties."  
  \item A Tier V character doesn’t need "a stronger Follower"—they need "a Follower with their own agenda, Obligations, and potential to betray."  
\end{itemize}
\textit{The challenge isn’t in the mechanical bonus—it’s in navigating the web of relationships and responsibilities they create.}
\end{tcolorbox}

\section{Common Pitfalls and Fixes}
\label{sec:pitfalls}
\index{assets!pitfalls}\index{followers!pitfalls}\index{GM!pitfalls}
\textbf{As GM, watch for these issues and apply these fixes:}  

\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Pitfall} & \textbf{Fix} \\
\midrule
\textbf{Asset hoarding:} Players collect Assets like XP, neglecting most. & Enforce Capacity (optional rule: Assets > Spirit cost +50\% upkeep or extra downtime scene). Make neglect hurt—advance timers, spend SB, or degrade benefits. \\
\midrule
\textbf{Follower automation:} Players treat Followers as scene-solvers. & Remind them: Followers only act through Bonds/Assists (Boon spend) or offscreen work. Never grant autonomous scene actions. \\
\midrule
\textbf{Power creep:} Assets/Followers make rolls too easy. & Never grant flat modifiers (+2 die). Always use narrative benefits (Position shift, DV $-1$, timer fill). If it feels too strong, add a cost (e.g., "Using the Safehouse tonight costs 1 Obligation to the guild"). \\
\midrule
\textbf{Neglect ignored:} Players overlook Asset/Follower decay. & Make neglect visible and consequential: describe the decay (rust, empty coffers, strained voices) and tie it to immediate complications (SB spends, timer advances). \\
\midrule
\textbf{One Asset/Follower dominance:} One investment overshadows all others. & Enforce spotlight spread: require players to rotate focus. If one Asset dominates, advance a neglect timer on it to force diversification. \\
\midrule
\textbf{Free Use abuse:} Players treat Free Use as automatic. & Only grant Free Use if fiction supports it. Require a Boon spend (Scene Surge) or downtime scene for offscreen benefits when neglected or unprepared. \\
\bottomrule
\end{tabularx}

\section{Optional Advanced Rules}
\label{sec:advanced-rules}
\index{assets!advanced}\index{followers!advanced}\index{GM!advanced}
Use these sparingly to add depth—but only when the fiction demands it.  

\subsection{Asset Networks (GM Guidance)}
\label{sec:asset-networks}
For players with multiple Assets, \textbf{as GM, allow them to declare a "Network" once per session} (cost: 1 Boon or 1 downtime scene) to:  
\begin{itemize}
  \item Treat two Minor Assets’ upkeep as waived this downtime (like the Market Front Standard Asset benefit).  
  \item Grant +1 die to a roll involving \textbf{both} Assets (e.g., using a Safehouse \emph{and} Informant to gather intelligence).  
  \item Mitigate 1 SB from a complication tied to \textit{either} Asset (e.g., "Your Informant warns of the landlord’s visit—spend the Boon to avoid the rent dispute").  
\end{itemize}
\textit{GM Guidance:} Limit to one Network declaration per session. Networks should feel earned—require players to describe how the Assets connect (e.g., "I pay the Informant to tip off the Safehouse landlord about incoming inspectors").

\subsection{Follower Management (GM Guidance)}
\label{sec:follower-management}
For players with high-Cap Followers (Cap 3+), \textbf{as GM, allow them to delegate Asset management} once per session (cost: 1 Boon or 1 downtime scene) to:  
\begin{itemize}
  \item Have the Follower tend up to \textbf{Cap} Assets (e.g., a Cap 3 Follower tends 3 Assets).  
  \item The Follower’s own upkeep (XP or downtime scene) \textbf{automatically covers} the upkeep of delegated Assets—no separate cost.  
  \item If the Follower’s Loyalty becomes Strained or Broken, \textbf{all delegated Assets immediately become Neglected}.  
\end{itemize}
\textit{GM Guidance:} This creates meaningful tension—players must choose between tending Followers directly or delegating Asset management. Neglecting the Follower risks losing \textit{multiple} Assets at once.

\subsection{Obligation Integration (GM Guidance)}
\label{sec:obligation-integration}
When Assets or Followers are tied to a faction (e.g., Temple Complex, Informant Ring), \textbf{as GM, use them to drive Obligation}:  
\begin{itemize}
  \item Leveraging an Asset against faction interests adds 1--2 Obligation segments (e.g., "Using the Banking Charter to freeze a rival’s account adds 2 Obligation to the bank").  
  \item Neglecting a Follower tied to a faction adds 1 Obligation segment per session neglected (e.g., "Ignoring your Informant’s pleas adds 1 Obligation to the spy network").  
  \item Tending a Follower’s welfare reduces Obligation by 1 segment per scene (e.g., "Spending downtime with your Informant clears 1 Obligation").  
\end{itemize}
\textit{GM Guidance:} Track this Obligation visibly—use it to create complications when full (e.g., "Your Obligation to the bank is full—spend 1 SB to advance the \emph{Bank Audit [6]} timer").  

\section{Quick Reference (GM)}
\label{sec:quick-ref-assets-followers}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Asset/Follower Quick Reference (GM),fonttitle=\bfseries,breakable]
\begin{itemize}
  \item \textbf{Asset Costs}: Minor 4 XP / Standard 8 XP / Major 12 XP / Tier III 16--20 XP.  
        \textit{GM Guidance:} Adjust slightly for narrative weight (e.g., a unique artifact might cost 6 XP as a Minor Asset).  
  \item \textbf{Follower Costs}: Cap $\times$ 2 XP (e.g., Cap 1 = 2 XP, Cap 3 = 6 XP, Cap 5 = 10 XP).  
        \textit{GM Guidance:} Adjust for story weight (e.g., a unique spirit ally might cost 3 XP as Cap 1).  
  \item \textbf{Asset Free Use}: Once per session, off-screen leverage \emph{only if fiction supports it}. Frame at scene start or as flashback with GM assent (cost: 1 Boon or 1 downtime scene).  
  \item \textbf{Asset Scene Surge}: Spend 1 Boon for a one-beat, on-screen intervention. \textbf{Limit to one meaningful use per scene}.  
  \item \textbf{Follower Benefits}: Only through Bonds (1 Boon for insight/support) or Assists (1 Boon for +1 die, max +3 total).  
  \item \textbf{Asset Upkeep}: Pay XP = base cost \emph{OR} devote a downtime scene. Neglect degrades condition (Maintained $\rightarrow$ Neglected $\rightarrow$ Compromised).  
  \item \textbf{Follower Upkeep}: Pay XP = Cap $\times$ 2 \emph{OR} devote a downtime scene tending them. Neglect degrades Loyalty/Fitness state.  
  \item \textbf{Asset Capacity (Optional)}: Assets $\leq$ Spirit are easy; beyond that, upkeep taxes mount (+50\% per extra Asset) or requires extra downtime scene.  
  \item \textbf{Follower Capacity (Optional)}: Followers $\leq$ Presence are easy; beyond that, upkeep taxes mount or requires extra downtime scene.  
  \item \textbf{Neglect is Your Engine}: Always use Asset/Follower neglect to create complications—advance timers, spend SB, or degrade benefits. \textbf{Never} let neglect be "nothing happens."  
\end{itemize}
\end{tcolorbox}

\section{Summary}
\label{sec:summary-assets-followers}
\textbf{As GM, remember:}  
\begin{itemize}
  \item Assets and Followers are \textbf{narrative levers}, not mechanical bonuses—they shape the world \emph{between scenes}, not \emph{during} them.  
  \item Their \textbf{true value lies in the hooks they create**—what happens if tended? what if neglected?**  
  \item \textbf{Neglect is your best storytelling tool**—use it to advance timers, spend SB, or degrade benefits.  
  \item \textbf{Always tie benefits to fiction**—never grant Free Use or Follower assistance automatically. Require a Boon spend, downtime scene, or narrative cost when unprepared or neglected.  
  \item \textbf{Spotlight spread is essential**—rotate which Assets/Followers take the stage so each investment feels earned and consequential.  
  \item \textbf{At high tiers, complexity > power**—the challenge isn’t in stronger bonuses, but in navigating webs of loyalty, obligation, and competing interests.  
\end{itemize}
Choose what to own and who to follow—then let the world answer. The best Assets and Followers create as many problems as they solve, and the best complications come from neglect, not power.  

\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title=From Faction to Asset/Follower: Taking Over a Terrestrial Patron (GM Guidance),label=sec:takeover-sidebar,breakable]
\index{terrestrial patrons!takeover}\index{assets!from factions}\index{followers!from factions}\index{GM!takeover}
Sometimes you stop owing the guild -- and start owning it. Rising to lead a faction (Thieves' Guild, Mercenary Company, Banking Charter, etc.) is a major story arc, not a simple XP purchase. \textbf{As GM, run this arc as a 2--3 session collaboration with the player.}  

\paragraph{Narrative Prerequisites (GM Verification)}
Before approving a takeover attempt, verify:  
\begin{itemize}
  \item The player has at least \textbf{6 Obligation} with the faction (deep entanglement).  
  \item They completed a \textbf{major service} for the faction (your discretion--e.g., saving it from destruction, winning a war, exposing a traitor at great cost).  
  \item They secured at least one \textbf{internal ally} with significant influence (another captain, a guildmaster, a wealthy backer).  
\end{itemize}
\textit{If prerequisites aren't met, the takeover attempt fails narratively--no XP cost, just a complication.}

\paragraph{The Takeover Arc (GM Collaboration)}
Collaborate on a 2--3 session arc that typically involves:  
\begin{enumerate}
  \item \textbf{Crisis or Power Vacuum} -- The current leader dies, disappears, or is disgraced (you define the crisis).  
  \item \textbf{Claim of Right} -- The player presents a valid claim (charter, election, duel victory, popular support--you judge validity).  
  \item \textbf{Consolidation} -- Defeat or co‑opt rivals, bribe or intimidate dissenters, and secure the faction's resources (you play the opposition).  
\end{itemize}
\textit{This arc should test the player's ingenuity, not just their XP pool.}

\paragraph{Mechanical Cost (GM Enforcement)}
Once the arc is complete, the player must spend \textbf{12 XP} (the cost of a Major Asset) to convert the faction into a personal Asset. \textbf{As GM:}  
\begin{itemize}
  \item Enforce this cost strictly--no conversions without payment.  
  \item Describe the conversion as institutional restructuring: "The ledgers are rewritten, the oaths renewed--it's yours now, but the work has only begun."  
\end{itemize}

\paragraph{Converting the Asset/Follower (GM Guidance)}
The new Asset uses the statistics of the most appropriate existing Asset from this compendium, modified by the faction's nature:  
\begin{itemize}
  \item \textbf{Thieves' Guild} $\rightarrow$ \textbf{Informant Ring} (Standard) or \textbf{Smuggler's Network} (Standard)  
  \item \textbf{Mercenary Company} $\rightarrow$ \textbf{Mercenary Contract} (Minor/Standard) or \textbf{Military Outpost} (Major)  
  \item \textbf{Banking Charter} $\rightarrow$ \textbf{Banking Charter} (Major) or \textbf{Trading Consortium} (Major)  
  \item \textbf{Royalist Network} $\rightarrow$ \textbf{Intelligence Bureau} (Specialized) or \textbf{Diplomatic Enclave} (Major)  
  \item \textbf{Ecktorian Mystery Cult} $\rightarrow$ \textbf{Temple Complex} (Standard) or \textbf{Magical Laboratory} (Specialized)  
\end{itemize}
\textit{As GM, feel free to create a custom Asset using \S\ref{sec:creating-your-own-asset} if none fit the faction's nature.}

\paragraph{What You Gain (GM Description)}
Upon conversion, grant:  
\begin{itemize}
  \item The Asset's normal \textbf{Free Use} and \textbf{Scene Surge} benefits.  
  \item The \textbf{Access Tag} associated with the faction (e.g., \textit{Guild-Ledgered}, \textit{Harbor Handsigns}).  
  \item A permanent \textbf{obligation track} toward the faction's members (now your underlings). This track is smaller (4 segments) and fills when you neglect their welfare or make reckless decisions that endanger them.  
\end{itemize}
\textit{As GM, track this obligation and use it to create complications when neglected (e.g., 'The crew grumbles--pay is late again. After 2 sessions neglected, they threaten to walk.')}

\paragraph{What You Lose (GM Enforcement)}
Enforce these losses strictly:  
\begin{itemize}
  \item The player can no longer use the faction as a \emph{terrestrial patron} to gain Obligation‑free perks--they are now responsible for it.  
  \item The faction becomes a target for your enemies--advance a \textbf{Rival Faction timer} whenever they use the Asset aggressively.  
  \item They must pay normal \textbf{upkeep} (XP or downtime scenes) to keep the Asset \emph{Maintained}. Neglect causes internal dissent, corruption, or defections--\textbf{use this as your complication engine}.  
\end{itemize}

\paragraph{Optional: Gradual Takeover (GM Permission)}
For a slower, more story‑driven approach, the player may forgo the 12 XP cost but must instead \textbf{spend a downtime scene per session} managing the faction (equivalent to Intensive upkeep) and accept that the Asset's benefits are unreliable until they have successfully completed a full arc of consolidation. \textbf{As GM, you may allow the XP cost to be paid in installments (e.g., 4 XP per downtime period) to represent gradual investment.}  

\paragraph{Example: Rellan Takes the Black Ledger (GM Narration)}
Rellan has served the Black Ledger smuggling syndicate for two arcs, accruing 6 Obligation. After exposing a rival who betrayed the syndicate (your narration), the aging leader steps down. Rellan spends 12 XP, completes a short arc consolidating loyal captains (you play the opposition), and converts the syndicate into a \textbf{Smuggler's Network (Standard Asset)}. He now commands hidden routes and safehouses--but must keep the crew paid and protected, or an internal revolt timer will start ticking.  
\textit{As GM, describe the takeover's cost: 'The syndicates' marks are scratched from the walls. New oaths are sworn--and new enemies are made.'}  
\end{tcolorbox}

\clearpage

\chapter*{Compendium of Assets}
\label{ch:compendium-assets}
\index{assets}\index{holdings}\index{influence}\index{GM!assets}

In \textit{Fate's Edge}, \textbf{Assets} are your off-screen levers---properties, charters, networks, and institutions that change the fiction between scenes. They do not fight beside you (see \S\ref{ch:assets-followers} for \emph{Followers}) but they shape the ground you walk on. As GM, your role is to:  
\begin{itemize}
  \item Introduce assets as narrative rewards (not automatic power-ups).  
  \item Price them appropriately to campaign tier and fiction.  
  \item Adjudicate their benefits \textbf{only when narratively justified}.  
  \item Use their neglect as a source of meaningful complications.  
\end{itemize}
Assets thrive when they create tension between player ambition and worldly consequences—not when they become spreadsheets.

\section{Design Principles for GM Use}
\index{assets!design principles}\index{GM!assets}
When introducing or homebrewing assets for your campaign, remember:  
\begin{itemize}
  \item \textbf{Assets shift fiction between scenes}: Use them to open doors, quiet rumors, stock larders, or calm crowds \emph{off-screen}. Their power lies in what happens \emph{when players aren't looking}.  
  \item \textbf{Assets create hooks}: Every asset should suggest \emph{both} opportunities (if tended) \emph{and} complications (if neglected). A neglected asset isn't "broken"—it's a story waiting to happen.  
  \item \textbf{Stay narrative-first}: Prefer position, scale, and access over flat modifiers. Instead of "+2 die," grant "start \textbf{Dominant} in negotiations with the guild."  
  \item \textbf{Never let assets replace player agency}: If an asset solves a problem too cleanly, add a cost (obligation, timer advance, or narrative cost).  
\end{itemize}

\paragraph{What Assets Are \emph{Not} (GM Guidance)}
As GM, ensure assets \emph{never}:  
\begin{itemize}
  \item Act autonomously \emph{inside} a scene (that's Followers' role).  
  \item Generate wealth or XP passively (their value is in narrative leverage, not bookkeeping).  
  \item Stack identical effects—if two assets would grant the same edge, the GM chooses the \emph{best narrative fit} (not the strongest mechanical bonus).  
\end{itemize}

\paragraph{Narrative First, Mechanics Second (GM Application)}
When an asset grants ``start \textbf{Dominant}'', it applies to social/informational rolls where positioning matters. For practical challenges (lock-picking, repairs), ``reduce DV by 1'' is preferred. \textbf{As GM, choose the form that best serves the fiction:}  
- If the asset represents \textit{influence} (e.g., a noble's favor), use Position shifts.  
- If it represents \textit{tools or preparation} (e.g., a well-stocked workshop), use DV reduction.  
- When in doubt, ask: ``Does this change \textit{how} the character approaches the problem, or just make the roll easier?'' Favor the former.

\section{Introducing Assets: GM Best Practices}
\index{assets!introduction}\index{GM!assets}
Assets should feel earned, not bought. Use these guidelines when placing assets in your campaign:  
\begin{itemize}
  \item \textbf{Tier Appropriateness}:  
        \begin{itemize}
          \item \textbf{Tier I (0--40 XP)}: Assets are rare and local (e.g., a single safehouse, a trusted informant). Cost 4 XP feels significant.  
          \item \textbf{Tier II (41--90 XP)}: Assets begin to show regional reach (e.g., a guild seat, a harbor office). Standard (8 XP) assets emerge as meaningful investments.  
          \item \textbf{Tier III (91--150 XP)}: Assets scale to national influence (e.g., a banking charter, a fortress). Major (12 XP) assets become campaign-defining.  
          \item \textbf{Tier IV+ (151+ XP)}: Assets are empire-shaping (see Tier III Institutions, \S\ref{sec:tier-iii-assets}). Reserve 16--20 XP assets for hard-won victories.  
        \end{itemize}
  \item \textbf{Fiction First, Cost Second}: Never assign an asset cost based solely on XP. Ask:  
        \begin{itemize}
          \item What narrative justification exists for this asset? (e.g., "The character saved the merchant's caravan—now they're granted a trade charter.")  
          \item Does the asset's scope match the fiction? (A village herbalist shouldn't grant a "National Healing House" asset.)  
          \item What complication arises if neglected? (This is your hook for future scenes.)  
        \end{itemize}
  \item \textbf{Price for Narrative Weight, Not Power}: A 4 XP asset in Tier I should feel as impactful as a 12 XP asset in Tier III—because it represents a larger \textit{proportion} of the character's influence. Adjust XP costs slightly up/down to reflect narrative significance (e.g., a unique artifact might cost 6 XP as a Minor Asset due to story weight).  
  \item \textbf{Always Attach a Hook}: When introducing an asset, immediately establish:  
        \begin{itemize}
          \item \textbf{What happens if tended?} (e.g., "The informant ring whispers of a shipment tonight.")  
          \item \textbf{What happens if neglected?} (e.g., "If ignored for two sessions, the informants go silent—or worse, start talking to rivals.")  
        \end{itemize}
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title={GM Tip: The Asset Neglect Engine},breakable]
\index{assets!neglect}
\textbf{Neglect is your best storytelling tool.} When players ignore an asset:  
\begin{itemize}
  \item Advance a relevant timer (e.g., \emph{Informant Network [4]} ticks +1 per session neglected).  
  \item Spend 1--2 SB to create a complication tied to the asset's decay (e.g., "The safehouse landlord demands rent—or eviction.").  
  \item Let the fiction degrade the asset's benefit (e.g., a neglected Scholar's Cell now requires a DV +1 roll for research).  
\end{itemize}
\textit{Never} treat neglect as "nothing happens." A neglected asset is a ticking time bomb—or a chance for players to re-engage with their world.
\end{tcolorbox}

\section{Asset Mechanics (GM Procedures)}
\label{sec:asset-mechanics}
\index{assets!mechanics}\index{upkeep}\index{Story Beats (SB)}
All assets share these core mechanics. \textbf{As GM, apply them consistently:}  
\begin{description}
  \item[Cost] Minor \textbf{4 XP}, Standard \textbf{8 XP}, Major \textbf{12 XP}. \index{assets!minor}\index{assets!standard}\index{assets!major}  
        \textit{GM Guidance:} Treat cost as a narrative threshold. In Tier I, 4 XP is a major investment; in Tier IV, it's pocket change. Adjust slightly for story weight (see above).  
  \item[Establishment] 1 day / 1 week / 1 month respectively to become operational.  
        \textit{GM Guidance:} Use this for downtime storytelling. "The safehouse takes a week to prepare—what do you do while waiting?"  
  \item[Upkeep] Each downtime, either pay XP equal to the Asset's base cost \emph{or} devote a full scene tending it. Neglect degrades condition (Maintained $\rightarrow$ Neglected $\rightarrow$ Compromised).  
        \textit{GM Guidance:} Track condition visibly. When an asset becomes Neglected, it's your cue to introduce complications. Compromised means it's broken—fixing it requires a quest or major downtime.  
  \item[Free Use] Once per session, trigger the listed \emph{off-screen} benefit (framed at scene start or as flashback with GM assent).  
        \textit{GM Guidance:} This is \textbf{not} automatic. Ask: ``Does the fiction support using this benefit \emph{now}?'' If yes, grant it; if no, require a scene of preparation or a Boon spend (Scene Surge).  
  \item[Scene Surge] Spend \textbf{1 Boon} to turn an Asset into a dramatic \emph{on-screen} intervention as described (one beat, fiction permitting).  
        \textit{GM Guidance:} This is the asset's "big moment." Limit to \textbf{one meaningful use per scene}—more than that feels like a deus ex machina. Describe it vividly (see \S\ref{sec:using-assets}).  
  \item[Extra Use] Spend \textbf{2 XP} to trigger the off-screen benefit an additional time this session.  
        \textit{GM Guidance:} Rarely allow this—it risks making assets feel like vending machines. Reserve for narrative crescendos (e.g., "You've poured everything into this safehouse—use it twice this session to flee the duke's guards.").  
  \item[Risk Interface] When an asset covers risk, it typically \textbf{improves Position}, \textbf{lowers DV by 1}, \textbf{fills 1 segment on a relevant timer}, or \textbf{mitigates 1 SB} (Story Beat) from a thematically linked complication.  
        \textit{GM Guidance:} This is where assets shine. Instead of "+2 die," say "Your contacts in the watch mean the patrol arrives 10 minutes late—start \textbf{Dominant}." Make the benefit feel earned by the fiction.  
\end{description}

\paragraph{Capacity (Optional GM Rule)}\index{assets!capacity}\index{GM!capacity}
A character comfortably maintains a number of assets up to their \textbf{Spirit} rating. For each asset beyond this:  
\begin{itemize}
  \item Increase its upkeep by +50\% (round up) \emph{OR}  
  \item Require an additional downtime scene to keep it \emph{Maintained}.  
\end{itemize}  
\textit{GM Guidance:} Use this to prevent asset hoarding. At Spirit 3, a 4th asset should feel like a strain—perhaps requiring a scene of frantic paperwork or causing minor complications until tended. This creates natural tension between ambition and upkeep.

\paragraph{Cross‑Reference to Faction Ties}\index{assets!faction ties}
Assets tied to a specific faction (e.g., Temple Complex, Banking Charter, Guild Seat) may generate \textbf{Obligation} if used recklessly or ignored. As GM, consult \S\ref{ch:world-powers-obligation} to:  
\begin{itemize}
  \item Award Obligation segments when players leverage assets against faction interests.  
  \item Let Obligation decay naturally if players tend the faction's welfare.  
  \item Use neglected assets as Obligation sources (e.g., "Ignoring the temple's pleas adds 1 Obligation to the gods each session neglected").  
\end{itemize}

\section{Minor Assets (4 XP): GM Tools for Local Leverage}
\label{sec:minor-assets}
\index{assets!minor}\index{GM!minor assets}
Minor assets are small holdings, charters, or kits that grant reliable, localized leverage. \textbf{As GM, use them to:}  
\begin{itemize}
  \item Reward early-game ingenuity without unbalancing the campaign.  
  \item Create intimate, character-driven story hooks (e.g., a family heirloom, a neighborhood favor).  
  \item Introduce complications that feel personal and immediate when neglected.  
\end{itemize}
Each entry below includes \textbf{GM-specific guidance} on introduction, adjudication, and neglect consequences.

\begin{description}
  \item[\textbf{Signature Weapon}] \index{signature weapon}  
        \emph{A named heirloom or masterpiece.}  
        \textbf{GM Free Use Guidance:} If gear damage would compromise it (per \S\ref{sec:equipment-condition}), downgrade to a minor flaw and require a quiet maintenance scene to clear—\emph{do not} auto-repair.  
        \textbf{GM Scene Surge Guidance:} Produce it at the perfect moment; improve Position one step for the first attack \emph{this scene}. Describe the draw: ``As the bandit lunges, your father's blade sings free—your grip is sure.''  
        \textbf{GM Neglect Hook:} If unused for 2+ sessions, the weapon requires repair (Cost: 1 XP or a downtime scene). Neglect further risks it being stolen or broken beyond repair.  
        \textit{Example:} ``The hilt is loose—swinging it feels like shaking a tooth. Repair it, or it flies from your grip next clash.''  

  \item[\textbf{Scholar's Cell}] \index{scholar's cell}  
        \emph{A quiet room, permits, and a custodian.}  
        \textbf{GM Free Use Guidance:} Begin one research scene at \textbf{Dominant} (or lower DV by 1). \emph{Only} grant this if the fiction supports it (e.g., the cell is prepared, the custodian is present).  
        \textbf{GM Scene Surge Guidance:} Reveal a cited source now; fill 1 segment on a relevant \emph{Information} timer. Describe the find: ``Between the mossy pages, you find the exact passage—ink still wet.''  
        \textbf{GM Neglect Hook:} If neglected, the custodian leaves or permits lapse. Research now requires a DV +1 roll (bribes, forgery, or breaking in). After 3 sessions neglected, the cell is seized by authorities.  
        \textit{Example:} ``The landlord knocks: 'Your rent's due—and I've heard whispers about those strange books.' Pay up, or they're calling the watch.''  

  \item[\textbf{Safehouse Network (Quarter)}] \index{safehouse}  
        \emph{Trapdoors, false walls, and a landlord who never saw you.}  
        \textbf{GM Free Use Guidance:} Start an entry/exit scene \textbf{Dominant}. \emph{Only} grant if the safehouse is prepared and stocked (track via downtime scenes).  
        \textbf{GM Scene Surge Guidance:} Produce a hidden egress; convert one \emph{pursuit} consequence into a temporary complication you can interact with. Describe the escape: ``Behind the wine rack—a stone slides aside. Cool night air hits your face as you slip into the alley.''  
        \textbf{GM Neglect Hook:} If neglected, the landlord grows suspicious or sells the property. After 1 session neglected, safehouse use costs 1 Boon to start Dominant; after 2 sessions, it's compromised (requires a scene to re-secure).  
        \textit{Example:} ``The floorboard creaks differently now. Your landlord eyes you over tea: 'Business has been slow. I wonder why.'''  

  \item[\textbf{Artisan's Bench}] \index{artisan's bench}  
        \emph{Tools for field repairs and fabrication.}  
        \textbf{GM Free Use Guidance:} Remove \emph{Compromised} from one mundane item between scenes. \emph{Only} grant if the bench is maintained and stocked (tracking via downtime).  
        \textbf{GM Scene Surge Guidance:} Improvise a single-use tool; gain a \emph{Boon} on the immediate Craft check. Describe the improvisation: ``With wire and prayer, you bend the lockpick just so—it slips in smooth.''  
        \textbf{GM Neglect Hook:} If neglected, tools rust or go missing. Repair requires a downtime scene + 1 XP. After 2 sessions neglected, the bench is Compromised (no conversions work until fixed).  
        \textit{Example:} ``The file's teeth are choked with rust. One good stroke—and it snaps. Time to visit the smith.''  

  \item[\textbf{Courier Chain}] \index{courier network}  
        \emph{Runners, dead-drops, and codes in one city.}  
        \textbf{GM Free Use Guidance:} Move a message or small parcel safely; start the related parley \textbf{Dominant}. \emph{Only} grant if the chain is active (track via downtime).  
        \textbf{GM Scene Surge Guidance:} A courier appears now; mitigate 1 SB tied to a missed appointment or lost message. Describe the arrival: ``A breathless kid skids to a halt, packet in hand: 'He said to give this to you—now.'''  
        \textbf{GM Neglect Hook:} If neglected, couriers go missing or codes change. After 1 session neglected, message delivery requires a Boon spend; after 2 sessions, the chain is Compromised (messages arrive late, altered, or not at all).  
        \textit{Example:} ``The drop-box is empty. The usual sign—a blue ribbon—is gone. Someone's been here.''  

  \item[\textbf{Small Holding (Farm/Orchard)}] \index{farm}\index{orchard}  
        \emph{Steady larder and a trusted foreman.}  
        \textbf{GM Free Use Guidance:} Waive a minor Supply complication or clear Fatigue 1 for one PC during downtime. \emph{Only} grant if the holding is tended (track via downtime scenes).  
        \textbf{GM Scene Surge Guidance:} Produce a timely wagon, mule, or crates to justify a plausible cover story. Describe the arrival: ``The creak of wheels on gravel—just in time. Three sacks of grain tumble from the cart as you dive behind it.''  
        \textbf{GM Neglect Hook:} If neglected, crops fail or foreman quits. After 1 session neglected, clear Fatigue 1 only once per downtime; after 2 sessions, the holding is Strained (yields half, requires 1 XP to maintain).  
        \textit{Example:} ``The orchard's quiet too quiet. No birds, no wind—and the trees are dropping fruit before it's ripe.''  

  \item[\textbf{Neighborhood Inn/Tavern}] \index{inn}\index{tavern}  
        \emph{A familiar house of bread and rumor.}  
        \textbf{GM Free Use Guidance:} First social opener in that neighborhood starts \textbf{Dominant}. \emph{Only} grant if the inn is maintained and the proprietor is friendly (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Call in the proprietor; shift a hostile crowd one step toward neutral. Describe the intervention: ``The innkeeper steps between you and the mob, arms wide: 'Not in my house. Take it outside—or buy a round.'''  
        \textbf{GM Neglect Hook:} If neglected, rumors turn sour or the inn changes hands. After 1 session neglected, social openers start at Controlled; after 2 sessions, the inn is Strained (proprietor drunk or surly—--1 die on social rolls there).  
        \textit{Example:} ``The usual smile is tight. 'Rough crowd tonight,' the proprietor mutters, wiping a glass that's already clean.''  

  \item[\textbf{Ritual Niche/Shrine}] \index{shrine}\index{ritual niche}  
        \emph{A consecrated corner with a small flame.}  
        \textbf{GM Free Use Guidance:} Clear Fatigue 1 or mitigate 1 SB from spiritual/psychological complications (fear, despair, doubt). \emph{Only} grant if the niche is maintained and consecrated (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Sanctify a doorway for one scene; hostile supernatural presence must reveal itself or back away one band (\S\ref{ch:world-interaction}). Describe the effect: ``You press your palm to the lintel—warmth flows like water. The shadow at the door flinches and retreats.''  
        \textbf{GM Neglect Hook:} If neglected, the flame sputters or the niche is desecrated. After 1 session neglected, mitigate only 0.5 SB (round down); after 2 sessions, it's Compromised (no spiritual benefits until reconsecrated).  
        \textit{Example:} ``The wick is low, the oil gone. The flame guttered out hours ago—leaving only cold stone and regret.''  

  \item[\textbf{River Skiff/Barge Share}] \index{boat}\index{barge}  
        \emph{Limited access to a hull and a taciturn pilot.}  
        \textbf{GM Free Use Guidance:} Water crossing scenes begin \textbf{Dominant}. \emph{Only} grant if the skiff is maintained and pilot is available (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Evade a patrol with a practiced channel; reduce DV by 1 on the escape roll. Describe the maneuver: ``You lean into the turn—current catches the keel, sliding you into the blind eddy where the patrol's lanterns can't reach.''  
        \textbf{GM Neglect Hook:} If neglected, the hull leaks or pilot is unavailable. After 1 session neglected, water crossings require a Boon spend to start Dominant; after 2 sessions, the share is Compromised (no access until repaired/replaced).  
        \textit{Example:} ``The pilot eyes the cracks in the hull. 'One more storm, and we're swimming. Pay for tar, or we walk.'''  

  \item[\textbf{Herbal Plot}] \index{herbs}\index{garden}  
        \emph{Drying racks, clay pots, and a steady hand at dawn.}  
        \textbf{GM Free Use Guidance:} Remove Fatigue 1 from one character or gain a \emph{Boon} to a single mundane healing attempt this downtime. \emph{Only} grant if the plot is tended (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Produce a rare poultice; convert a Severe healing timer into two Moderate ticks. Describe the find: ``Under the damp cloth—a bundle of silverleaf, still glistening with dew.''  
        \textbf{GM Neglect Hook:} If neglected, herbs fail or plot is overgrown. After 1 session neglected, remove Fatigue 1 only once per downtime; after 2 sessions, it's Compromised (no herbs until replanted).  
        \textit{Example:} ``The mint's gone to seed. The soil's hard as baked clay—and the rabbits have turned it into a warren.''  

  \item[\textbf{Personal Library}] \index{library!personal}  
        \emph{Shelves of annotated tomes and a reading chair.}  
        \textbf{GM Free Use Guidance:} Gain a \emph{Boon} on one Lore roll per session. \emph{Only} grant if the library is maintained and accessible (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Quote a relevant passage; reduce DV of a related Investigation by 1. Describe the insight: ``You trace a finger along the margin—'Ah. So that's why the ritual failed.'''  
        \textbf{GM Neglect Hook:} If neglected, books are damaged or stolen. After 1 session neglected, gain a Boon only every other session; after 2 sessions, the library is Strained (--1 die on Lore rolls there until repaired).  
        \textit{Example:} ``The spine's cracked, the pages water-stained. One more rain—and that treaty copy is pulp.''  

  \item[\textbf{Stable/Coach House}] \index{stable}\index{transport}  
        \emph{Reliable mounts and drivers who ask few questions.}  
        \textbf{GM Free Use Guidance:} Land travel scenes begin \textbf{Dominant}. \emph{Only} grant if the stable is maintained and mounts/vehicles are ready (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Produce a fresh mount or spare vehicle; ignore the first \emph{exhaustion} complication. Describe the arrival: ``The stable boy leads out a chestnut—nostrils flared, coat gleaming. 'She's rested and ready,' he says.''  
        \textbf{GM Neglect Hook:} If neglected, mounts sicken or vehicles break down. After 1 session neglected, ignore exhaustion only once per downtime; after 2 sessions, the stable is Strained (--1 die on travel rolls there until repaired).  
        \textit{Example:} ``The mare's favoring her left foreleg. The vet's bill is steep—but better than a lame horse on the road.''  

  \item[\textbf{Alchemist's Workshop}] \index{alchemist}  
        \emph{Reagents, retorts, and a suspicious chimney.}  
        \textbf{GM Free Use Guidance:} Produce one dose of a common alchemical item between scenes. \emph{Only} grant if the workshop is maintained and stocked (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Brew a specialized compound; gain a \emph{Boon} on the next Craft (Alchemy) roll. Describe the brew: ``The liquid shifts from sickly green to deep blue—as if the sky decided to join the potion.''  
        \textbf{GM Neglect Hook:} If neglected, reagents degrade or equipment fails. After 1 session neglected, yield halved; after 2 sessions, Compromised (no production until restocked/repaired).  
        \textit{Example:} ``The vitriol's eaten through the glass. Time for new bottles—and a stern word to the supplier.''  

  \item[\textbf{Messenger Birds}] \index{messenger birds}\index{communication}  
        \emph{Trained pigeons or ravens with coded leg bands.}  
        \textbf{GM Free Use Guidance:} Send a simple message across moderate distance safely. \emph{Only} grant if the birds are maintained and trained (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Receive urgent news now; mitigate 1 SB from \emph{miscommunication}. Describe the arrival: ``A raven lands on your shoulder, legs twitching—'Danger. East gate. Now.'''  
        \textbf{GM Neglect Hook:} If neglected, birds go astray or codes break. After 1 session neglected, message sending requires a Boon spend; after 2 sessions, the network is Compromised (messages arrive late, wrong, or not at all).  
        \textit{Example:} ``The coop's silent. No wings beat the dusk—and the leg bands are dull with dirt. Someone forgot to feed them.''  
\end{description}

\section{Standard Assets (8 XP): GM Tools for Regional Influence}
\label{sec:standard-assets}
\index{assets!standard}\index{GM!standard assets}
Standard assets are established holdings with regional reach. \textbf{As GM, use them to:}  
\begin{itemize}
  \item Reward mid-tier accomplishments with meaningful influence (e.g., a guild seat, a harbor office).  
  \item Create hooks that tie into regional politics, trade, or faction play.  
  \item Introduce complications that scale with neglect (e.g., losing favor, inviting rivals).  
\end{itemize}
Each entry below includes \textbf{GM-specific guidance} on introduction, adjudication, and neglect consequences.

\begin{description}
  \item[\textbf{Charter or Minor Title}] \index{charter}\index{title}  
        \emph{Recognized papers and privileges.}  
        \textbf{GM Free Use Guidance:} Secure an audience or official hearing; start \textbf{Dominant}. \emph{Only} grant if the charter is valid and recognized (track via downtime/faction standing).  
        \textbf{GM Scene Surge Guidance:} Invoke prerogative to delay immediate sanctions; mitigate 1 SB of civic backlash. Describe the delay: ``You slap the seal on the table—'By order of the Marshal, this inquiry pauses till dawn.' The clerk pales and nods.''  
        \textbf{GM Neglect Hook:} If neglected, the charter lapses or is challenged. After 1 session neglected, secure audience starts at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to renew/defend).  
        \textit{Example:} ``The ink's faded, the seal cracked. The clerk squints: 'This expired last month. Renew it—or we start from scratch.'''  

  \item[\textbf{Guild Seat}] \index{guild}  
        \emph{Voting rights and shop‑rights.}  
        \textbf{GM Free Use Guidance:} Commission or procurement opens \textbf{Dominant} or DV $-1$. \emph{Only} grant if the seat is active and voting rights intact (track via faction standing).  
        \textbf{GM Scene Surge Guidance:} A guild factor vouches; fill 1 segment on a \emph{Permit/Access} timer. Describe the voucher: ``The factor slides a coin across the table—'For the Brotherhood. Consider it done.'''  
        \textbf{GM Neglect Hook:} If neglected, the seat is challenged or voting rights suspended. After 1 session neglected, commission starts at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to defend seat/rights).  
        \textit{Example:} ``The Master frowns at your slate. 'Your dues are three months late. Pay up—or we vote you out.'''  

  \item[\textbf{Market Front}] \index{market}\index{stall}  
        \emph{A fixed stall and regular customers.}  
        \textbf{GM Free Use Guidance:} Treat one \emph{Minor} Asset's upkeep as waived this downtime. \emph{Only} grant if the stall is operational and customers loyal (track via downtime/faction standing).  
        \textbf{GM Scene Surge Guidance:} Liquidate stock; justify on-hand mundane gear within reason. Describe the liquidation: ``You heap the counter with goods—'Everything must go!'' Coins clink as you count the takings.''  
        \textbf{GM Neglect Hook:} If neglected, customers dwindle or stock spoils. After 1 session neglected, waive upkeep only every other downtime; after 2 sessions, the front is Strained (requires 1 XP or a scene to revive trade).  
        \textit{Example:} ``The usual faces are missing. The bread's stale—and the fishmonger eyes your empty stall with pity.''  

  \item[\textbf{Informant Ring}] \index{spies}\index{informants}  
        \emph{Eyes and ears on retainer.}  
        \textbf{GM Free Use Guidance:} Targeted inquiry begins \textbf{Dominant}. \emph{Only} grant if the ring is active and informants are loyal (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Reveal a hidden schedule or route now; mitigate 1 SB from \emph{ambush/surprise}. Describe the reveal: ``A scrap of paper slips into your palm—'The convoy leaves at midnight. Take the north ridge.'''  
        \textbf{GM Neglect Hook:} If neglected, informants go silent or turn hostile. After 1 session neglected, targeted inquiry starts at Controlled; after 2 sessions, it's Strained (--1 die on inquiry rolls there until trust is rebuilt).  
        \textit{Example:} ``The drop-box is empty. The usual sign—a blue ribbon—is gone. Someone's been here.''  

  \item[\textbf{Mercenary Contract (Cap 2)}] \index{mercenaries}  
        \emph{Access to a small trained unit.}  
        \textbf{GM Free Use Guidance:} Introduce temporary \emph{off-screen} security that downgrades ``raid'' to ``attempted raid.'' \emph{Only} grant if the contract is active and mercenaries are paid (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} One on-screen intervention (one beat) that improves Position for a withdrawal or breach. Describe the intervention: ``Horns sound from the ridge—mercenaries pour over the crest, shields locking. The raiders hesitate.''  
        \textbf{GM Neglect Hook:} If neglected, mercenaries leave or demand higher pay. After 1 session neglected, security downgrade requires 1 Boon spend; after 2 sessions, the contract is Strained (unit reduced or unreliable until paid/replaced).  
        \textit{Example:} ``The captain spits in the dirt. 'Pay's late. We'll hold the line—but don't count on us for the counter.'''  

  \item[\textbf{Trading Charter}] \index{trade}\index{charter!trading}  
        \emph{Right-of-way at gates and tolls.}  
        \textbf{GM Free Use Guidance:} Border/inspection scenes start \textbf{Dominant}. \emph{Only} grant if the charter is valid and honored (track via downtime/faction standing).  
        \textbf{GM Scene Surge Guidance:} Produce stamped manifests; DV $-1$ on the next scrutiny roll. Describe the manifest: ``You unfurl the documents—seals intact, weights exact. The guard nods and waves you through.''  
        \textbf{GM Neglect Hook:} If neglected, the charter lapses or tolls increase. After 1 session neglected, border scenes start at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to renew/defend rights).  
        \textit{Example:} ``The clerk frowns at your papers. 'Tolls went up yesterday. Pay the difference—or turn back.'''  

  \item[\textbf{Licensed Library}] \index{library}  
        \emph{Stacks and a learned custodian.}  
        \textbf{GM Free Use Guidance:} Deep‑topic research begins \textbf{Dominant} and gains a \emph{Boon}. \emph{Only} grant if the library is licensed and custodian is present (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Cite a precedent to force a pause in debate (one exchange). Describe the citation: ``You slam the tome shut—'Precedent 7B. The chair must yield.' The hall falls silent.''  
        \textbf{GM Neglect Hook:} If neglected, the license lapses or custodian leaves. After 1 session neglected, deep research starts at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to renew license/replace custodian).  
        \textit{Example:} ``The custodian packs her satchel. 'The Bishop's men came for the license. I'm heading to the abbey—good luck.'''  

  \item[\textbf{Craftsman's Quarter}] \index{workshops}\index{apprentices}  
        \emph{Shop plus apprentices.}  
        \textbf{GM Free Use Guidance:} Clear \emph{Compromised} from a crafted item or structure. \emph{Only} grant if the quarter is maintained and stocked (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Rush a bespoke piece; fill 1 segment on a \emph{Build/Repair} timer. Describe the rush: ``The forge blazes as you hammer—sparks flying like angry stars. By dawn, it's done.''  
        \textbf{GM Neglect Hook:} If neglected, apprentices leave or tools degrade. After 1 session neglected, clear Compromised only once per downtime; after 2 sessions, the quarter is Strained (requires 1 XP or a scene to restore staff/tools).  
        \textit{Example:} ``The anvil's cracked. One good swing—and it sends shards flying. Time for a new one—and a lecture on metallurgy.''  

  \item[\textbf{Temple Complex}] \index{temple}  
        \emph{Courts, choirs, charity.}  
        \textbf{GM Free Use Guidance:} Among adherents, start \textbf{Dominant} or mitigate 1 SB from \emph{crowd/morale}. \emph{Only} grant if the temple is maintained and active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Sanctuary: hostile social pressure must escalate openly or wait. Describe the sanctuary: ``You step into the hush—voices drop, eyes avert. For now, you are safe here.''  
        \textbf{GM Neglect Hook:} If neglected, adherents dwindle or charity falters. After 1 session neglected, start Dominant only every other session; after 2 sessions, it's Strained (requires 1 XP or a scene to restore trust/activity).  
        \textit{Example:} ``The collection plate is light. The priest's smile is tight—'Hard times, my child. But the gods still watch.'''  

  \item[\textbf{Harbor Office}] \index{harbor}\index{port}  
        \emph{Berths, booms, bells.}  
        \textbf{GM Free Use Guidance:} Priority loading/unloading; start \textbf{Dominant}. \emph{Only} grant if the office is maintained and honors are in place (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Hold the chain for one beat; negate a single \emph{pursuit by water} consequence. Describe the hold: ``You grip the chain lever—'Hold!'' The great iron links groan into place, blocking the channel.''  
        \textbf{GM Neglect Hook:} If neglected, berths silt up or honors lapse. After 1 session neglected, priority loading starts at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to dredge/renew honors).  
        \textit{Example:} ``The harbormaster checks the tide chart. 'Silt's built up at Berth 3. Pay for dredging—or we wait for the spring flood.'''  

  \item[\textbf{Healing House}] \index{healing}\index{infirmary}  
        \emph{Beds, herbs, bells.}  
        \textbf{GM Free Use Guidance:} During downtime, clear \emph{Harm 1} or \emph{Fatigue 2} from one ally. \emph{Only} grant if the house is maintained and stocked (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Stabilize now; convert a Severe injury consequence into a 4‑segment Recovery timer. Describe the stabilization: ``The healer nods—'She'll live. But she needs rest, clean bandages, and time.'''  
        \textbf{GM Neglect Hook:} If neglected, beds go bad or herbs spoil. After 1 session neglected, clear Harm 1/Fatigue 2 only once per downtime; after 2 sessions, the house is Strained (requires 1 XP or a scene to restock/repair).  
        \textit{Example:} ``The bells are silent. The wards are faded—and the cough in Ward 2 sounds wetter every day.''  

  \item[\textbf{Observatory}] \index{observatory}\index{navigation}  
        \emph{Lenses and ephemerides.}  
        \textbf{GM Free Use Guidance:} Predict timing; navigation/timing openers start \textbf{Dominant}. \emph{Only} grant if the observatory is maintained and instruments are calibrated (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Name a narrow window; DV $-1$ for an action attempted inside it. Describe the window: ``You adjust the astrolabe—'The eclipse peaks at 3:07 AM. That's your moment.'''  
        \textbf{GM Neglect Hook:} If neglected, instruments drift or lenses crack. After 1 session neglected, predictions require a Boon spend to start Dominant; after 2 sessions, it's Strained (requires 1 XP or a scene to recalibrate/replace).  
        \textit{Example:} ``The glass is spiderwebbed. The stars are hard to read—and the ephemeris is missing three months.''  

  \item[\textbf{Performance Venue}] \index{theater}\index{performance}  
        \emph{Stage and regulars.}  
        \textbf{GM Free Use Guidance:} Begin Performance with a receptive crowd. \emph{Only} grant if the venue is maintained and crowd is receptive (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Spotlight an ally; shift a crowd one step toward your desired mood. Describe the spotlight: ``You nod to the fiddler—she lifts her bow. The melody swirls, pulling eyes to your friend as they step forward.''  
        \textbf{GM Neglect Hook:} If neglected, the stage decays or crowd turns hostile. After 1 session neglected, begin Performance at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to repair/rebook).  
        \textit{Example:} ``The seats are squeaky. The curtain sticks halfway—and the usual crowd's been replaced by sailors looking for a fight.''  

  \item[\textbf{Hospitality House}] \index{hospitality}\index{guest right}  
        \emph{Beds, bread, rumor.}  
        \textbf{GM Free Use Guidance:} Gain a \emph{Protected pause} in parley on your ground. \emph{Only} grant if the house is maintained and guest-right is honored (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Invoke guest-right; hostile social pressure must step into the open. Describe the invocation: ``You rise, voice clear—'By the ancient right of hospitality, we speak in peace.' The hall stills, waiting.''  
        \textbf{GM Neglect Hook:} If neglected, beds go bad or rumor turns sour. After 1 session neglected, Protected pause requires a Boon spend; after 2 sessions, the house is Strained (requires 1 XP or a scene to restock/reinvite).  
        \textit{Example:} ``The loaves are dense as stone. The gossip's turned vicious—and the children won't come inside anymore.''  

  \item[\textbf{Storage Vault}] \index{vault}\index{storage}  
        \emph{Locks and ledgers, damp stone and iron.}  
        \textbf{GM Free Use Guidance:} Ignore the first \emph{theft/sabotage} targeting stored goods. \emph{Only} grant if the vault is maintained and secure (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Produce a cached item plausibly stored here. Describe the cache: ``You brush aside the dust—there it is, wrapped in oilcloth. The blade still gleams.''  
        \textbf{GM Neglect Hook:} If neglected, locks rust or ledgers rot. After 1 session neglected, ignore theft only every other downtime; after 2 sessions, the vault is Strained (requires 1 XP or a scene to repair/reinforce).  
        \textit{Example:} ``The ledger's pages are stuck together. One tug—and they tear. Time for new paper—and a dry cellar.''  

  \item[\textbf{Hidden Cache}] \index{cache}\index{emergency}  
        \emph{A buried lockbox or hollow tree, accessible anywhere in the city.}  
        \textbf{GM Free Use Guidance:} Once per session, access emergency gear when cornered; start the scene \textbf{Dominant}. \emph{Only} grant if the cache is maintained and accessible (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Reveal a larger stash; fill 1 segment on a \emph{Supply} timer. Describe the reveal: ``You pull aside the loose stone—there it is, just as you left it. The sack feels heavy with promise.''  
        \textbf{GM Neglect Hook:} If neglected, the cache is discovered or compromised. After 1 session neglected, access requires a Boon spend to start Dominant; after 2 sessions, it's Strained (requires 1 XP or a scene to re-hide/replenish).  
        \textit{Example:} ``The moss is disturbed. The hollow smells of damp earth—and the lockbox is gone. Someone's been here.''  

  \item[\textbf{Signal Relay}] \index{relay}\index{communications}  
        \emph{Fires, flags, drums.}  
        \textbf{GM Free Use Guidance:} Coordinated actions across distance start \textbf{Dominant}. \emph{Only} grant if the relay is maintained and signals are clear (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Push a ready signal now; mitigate 1 SB from \emph{confusion/lag}. Describe the push: ``You strike the drum—three sharp beats. The signal flies down the valley, clear and true.''  
        \textbf{GM Neglect Hook:} If neglected, signals fail or equipment degrades. After 1 session neglected, coordinated actions require a Boon spend to start Dominant; after 2 sessions, it's Strained (requires 1 XP or a session to repair/replace).  
        \textit{Example:} ``The drumskin's split. The fire's low—and the flags are rotting in the rain. Time for new hides and a steady hand.''  

  \item[\textbf{Merchant Caravan}] \index{caravan}\index{trade}  
        \emph{One wagon, two guards, and a route.}  
        \textbf{GM Free Use Guidance:} Trade missions begin \textbf{Dominant}. \emph{Only} grant if the caravan is maintained and route is secure (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Call in the caravan early; provide unexpected supplies or reinforcements. Describe the arrival: ``Hooves pound the ridge—then the wagon appears, creaking under sacks of grain. 'We brought extra,' the guard grunts.''  
        \textbf{GM Neglect Hook:} If neglected, the caravan disbands or route becomes unsafe. After 1 session neglected, trade missions require a Boon spend to start Dominant; after 2 sessions, it's Strained (requires 1 XP or a scene to re-form/re-secure route).  
        \textit{Example:} ``The guards eye the broken bridge. 'We can't cross here. Pay for a detour—or we turn back.'''  

  \item[\textbf{Legal Practice}] \index{law}\index{barrister}  
        \emph{Papers, precedents, and clients.}  
        \textbf{GM Free Use Guidance:} Legal proceedings start \textbf{Dominant} or DV $-1$. \emph{Only} grant if the practice is maintained and clients are active (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Produce a forgotten writ; negate one \emph{legal} complication. Describe the writ: ``You blow the dust off the parchment—'Ah. Here's the clause they missed.'''  
        \textbf{GM Neglect Hook:} If neglected, papers decay or clients leave. After 1 session neglected, legal proceedings start at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to restore papers/clients).  
        \textit{Example:} ``The ink's bled through the paper. One more flood—and that precedent is soup.''  

  \item[\textbf{Smuggler's Network}] \index{smugglers}\index{underground}  
        \emph{Contacts and hidden routes.}  
        \textbf{GM Free Use Guidance:} Illicit transport begins \textbf{Dominant}. \emph{Only} grant if the network is maintained and contacts are active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Bypass a checkpoint; mitigate 1 SB from \emph{customs/patrols}. Describe the bypass: ``You slip through the reeds where the fence is rusted through—just wide enough for the skiff.''  
        \textbf{GM Neglect Hook:} If neglected, contacts go silent or routes are compromised. After 1 session neglected, illicit transport starts at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to restore contacts/routes).  
        \textit{Example:} ``The drop-point is cold. The usual signal—a tapped stone—is missing. Someone's been here.''  

  \item[\textbf{Crafting Guildhall}] \index{guildhall}\index{crafting}  
        \emph{Tools, masters, and apprentices.}  
        \textbf{GM Free Use Guidance:} Complex crafting projects gain a \emph{Boon}. \emph{Only} grant if the guildhall is maintained and stocked (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Rush production; fill 1 segment on a \emph{Manufacturing} timer. Describe the rush: ``The looms fly as you shout—thread snapping like gunfire. By noon, it's done.''  
        \textbf{GM Neglect Hook:} If neglected, masters leave or tools degrade. After 1 session neglected, complex projects gain a Boon only every other session; after 2 sessions, it's Strained (requires 1 XP or a scene to restore masters/tools).  
        \textit{Example:} ``The spindle's worn smooth. One more pull—and it snaps. Time for new wire—and a stern look at the apprentice.''  
\end{description}

\section{Major Assets (12 XP): GM Tools for Regional Power}
\label{sec:major-assets}
\index{assets!major}\index{GM!major assets}
Major assets are significant holdings that bend regional outcomes. \textbf{As GM, use them to:}  
\begin{itemize}
  \item Reward high-tier accomplishments with transformative influence (e.g., a fortress, a banking charter).  
  \item Create hooks that tie into national politics, economics, or major faction play.  
  \item Introduce complications that scale with neglect (e.g., invasion, embargo, revolt).  
\end{itemize}
Each entry below includes \textbf{GM-specific guidance} on introduction, adjudication, and neglect consequences.

\begin{description}
  \item[\textbf{Ward License}] \index{city ward}\index{license}  
        \emph{You run a city district's levers.}  
        \textbf{GM Free Use Guidance:} In-ward civic openers start \textbf{Dominant}. \emph{Only} grant if the license is active and honors are in place (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Re-route patrols or light---shift a chase route 1 step in your favor. Describe the reroute: ``You nod to the lamplighter—'Extinguish the third lamp.' The light winks out, leaving the alley in shadow.''  
        \textbf{GM Neglect Hook:} If neglected, the ward lapses or honors are challenged. After 1 session neglected, civic openers start at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to renew license/defend honors).  
        \textit{Example:} ``The Warden frowns at your papers. 'Your tribute's late. Pay up—or we reassess the ward's boundaries.'''  

  \item[\textbf{Exclusive Right (Monopoly)}] \index{monopoly}  
        \emph{Control of a staple (salt, timber, ferries).}  
        \textbf{GM Free Use Guidance:} In market scenes, ignore the first \emph{supply} complication. \emph{Only} grant if the monopoly is active and honored (track via downtime/faction standing/Obligation).  
        \textbf{GM Scene Surge Guidance:} Declare a brief scarcity/surplus to leverage terms; DV $-1$ on a single negotiation. Describe the declaration: ``You slam your fist on the table—'The salt's mine till sundown. Name your price.'''  
        \textbf{GM Neglect Hook:} If neglected, the monopoly lapses or is challenged. After 1 session neglected, ignore supply complication only every other session; after 2 sessions, it's Strained (requires 1 XP or a scene to renew/defend rights).  
        \textit{Example:} ``The customs officer checks your seal. 'The monopoly expired yesterday. Pay the market rate—or we seize the shipment.'''  

  \item[\textbf{Fortress/Stronghold}] \index{fortress}\index{stronghold}  
        \emph{Stone, stores, signals.}  
        \textbf{GM Free Use Guidance:} In raid/siege framing, negate the first \emph{breach} consequence. \emph{Only} grant if the fortress is maintained and garrisoned (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Sound beacons; allies arrive in time to improve Position for withdrawal or stand. Describe the beacons: ``Three sharp blasts on the horn—answer comes from the ridge: 'We're coming. Hold the line.'''  
        \textbf{GM Neglect Hook:} If neglected, the fortress decays or garrison mutinies. After 1 session neglected, neglect breach consequence only every other session; after 2 sessions, it's Strained (requires 1 XP or a scene to repair/reinforce/garrison).  
        \textit{Example:} ``The mortar's crumbling in the northwest tower. One good storm—and we're digging bodies out of the rubble.'''  

  \item[\textbf{Demesne (Estate)}] \index{estate}\index{demesne}  
        \emph{Fields, mills, tenants.}  
        \textbf{GM Free Use Guidance:} Waive minor Supply fallout this arc or treat one additional Minor Asset as having free upkeep. \emph{Only} grant if the demesne is maintained and productive (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Muster teamsters and hands; fill 1 segment on a \emph{Logistics} timer. Describe the muster: ``You whistle—boots pound the dirt as wagons roll out. 'Move!'"  
        \textbf{GM Neglect Hook:} If neglected, fields lie fallow or mills seize up. After 1 session neglected, waive Supply fallout only every other arc; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to restore productivity).  
        \textit{Example:} ``The well's run dry. The mill's seized—and the tenants are whispering about leaving.'''  

  \item[\textbf{Fleet Share}] \index{fleet}  
        \emph{A cut of hulls and signals.}  
        \textbf{GM Free Use Guidance:} Maritime travel starts \textbf{Dominant}. \emph{Only} grant if the fleet is maintained and signals are honored (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Signal a safe corridor; mitigate 1 SB from \emph{storm/patrol}. Describe the signal: ``You raise the flag—'Clear passage through the Straits.' The pilot nods and eases the throttle.''  
        \textbf{GM Neglect Hook:} If neglected, the fleet deteriorates or signals lapse. After 1 session neglected, maritime travel starts at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to repair/reinforce/renew signals).  
        \textit{Example:} ``The bosun eyes the frayed ropes. 'One more gale—and we're swimming. Pay for new line—or we head for port.'''  

  \item[\textbf{College (University Seat)}] \index{university}\index{college}  
        \emph{Faculty, quad, seal.}  
        \textbf{GM Free Use Guidance:} In downtime, reduce \emph{one} skill advancement cost by 1 XP (minimum 1). \emph{Only} grant if the college is maintained and the seat is active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Convene a colloquy to force a pause in magical proceedings (\S\ref{ch:magic}). Describe the convening: ``You strike the gavel—'The Council is in session. All rise.'''  
        \textbf{GM Neglect Hook:} If neglected, faculty departs or seal is challenged. After 1 session neglected, reduce skill cost only every other downtime; after 2 sessions, it's Strained (requires 1 XP or a scene to renew seat/replace faculty).  
        \textit{Example:} ``The Dean signs the transfer. 'Your funding's cut. We'll teach what we can—but no new arts.'''  

  \item[\textbf{Banking Charter}] \index{bank}\index{letters of credit}  
        \emph{Letters, liens, a counting house.}  
        \textbf{GM Free Use Guidance:} Present credit to bypass petty fees and minor delays. \emph{Only} grant if the charter is maintained and honored (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Freeze a rival's access for one exchange; mitigate 1 SB from \emph{bribery/extortion}. Describe the freeze: ``You slide the order across the table—'Freeze Account 7B. Effective immediately.'''  
        \textbf{GM Neglect Hook:} If neglected, the charter lapses or is challenged. After 1 session neglected, present credit only every other session; after 2 sessions, it's Strained (requires 1 XP or a scene to renew/defend rights).  
        \textit{Example:} ``The teller checks the ledger. 'Your credit line's frozen. Pay the fee—or we close the account.'''  

  \item[\textbf{Diplomatic Enclave}] \index{diplomacy}\index{enclave}  
        \emph{Flag, immunity, ink.}  
        \textbf{GM Free Use Guidance:} In foreign courts, begin \textbf{Dominant}. \emph{Only} grant if the enclave is maintained and immunity is honored (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Invoke safe-conduct; hostile action must escalate openly or withdraw. Describe the invocation: ``You unfurl the treaty—'By the Accord of Veridian, we speak as equals.' The foreign envoy blinks and nods.''  
        \textbf{GM Neglect Hook:} If neglected, immunity lapses or enclave is compromised. After 1 session neglected, foreign court openers start at Controlled; after 2 sessions, it's Strained (requires 1 XP or a scene to renew immunity/rebuild trust).  
        \textit{Example:} ``The ambassador folds her arms. 'Your immunity's revoked. We'll talk—but only in the chancery.'''  

  \item[\textbf{Mine/Quarry}] \index{mine}\index{quarry}  
        \emph{Ore, stone, skilled crews.}  
        \textbf{GM Free Use Guidance:} Clear \emph{Compromised} on one structure or armament. \emph{Only} grant if the mine is maintained and crew is skilled (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Deliver timbers/stone now; DV $-1$ on a \emph{reinforce/repair} roll. Describe the delivery: ``The carts rumble out—stone piled high, timber stacked neat. 'Here's your order,' the foreman grunts.''  
        \textbf{GM Neglect Hook:} If neglected, ore veins fade or crews desert. After 1 session neglected, deliver timbers/stone only every other session; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to restore ore/crews).  
        \textit{Example:} ``The foreman spits in the dust. 'The vein's pinching out. Pay for a new shaft—or we abandon the claim.'''  

  \item[\textbf{Mining Operation}] \index{mining}\index{resources}  
        \emph{Shafts, carts, and crews.}  
        \textbf{GM Free Use Guidance:} Ignore the first \emph{resource shortage} complication this arc. \emph{Only} grant if the operation is maintained and crews are active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Strike a rich vein; produce unexpected wealth or materials. Describe the strike: ``The pick hits bedrock—then sings. 'Gold!' the miner whoops, holding up the nugget.''  
        \textbf{GM Neglect Hook:} If neglected, shafts flood or crews leave. After 1 session neglected, ignore resource shortage only every other arc; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to dewater/rehire crews).  
        \textit{Example:} ``The pumps are silent. The water's rising—and the foreman's face is grim. Time for new hands—or we admit defeat.'''  

  \item[\textbf{Military Outpost}] \index{military}\index{outpost}  
        \emph{Barracks, stores, and command.}  
        \textbf{GM Free Use Guidance:} Military operations begin \textbf{Dominant}. \emph{Only} grant if the outpost is maintained and garrisoned (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Call for reinforcements; improve Position for a tactical withdrawal or assault. Describe the call: ``You raise the horn—three long notes. The answer comes drifting over the hills: 'On the way. Hold for ten.'''  
        \textbf{GM Neglect Hook:} If neglected, the outpost decays or garrison mutinies. After 1 session neglected, call for reinforcements only every other session; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to repair/reinforce/rehire garrison).  
        \textit{Example:} ``The sergeant eyes the empty barracks. 'Half the squad's gone AWOL. We'll hold—but don't expect miracles.'''  

  \item[\textbf{Trading Consortium}] \index{consortium}\index{commerce}  
        \emph{Multiple fronts and shared capital.}  
        \textbf{GM Free Use Guidance:} Ignore the first price gouging or supply shortage complication this arc. \emph{Only} grant if the consortium is maintained and fronts are active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Coordinate a multi-front deal; DV $-1$ on a major commercial negotiation. Describe the coordination: ``You slide the contract across the table—'Signed, sealed, delivered. Let's get to work.'''  
        \textbf{GM Neglect Hook:} If neglected, fronts withdraw or capital dries up. After 1 session neglected, ignore price gouging only every other arc; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to restore fronts/capital).  
        \textit{Example:} ``The consul shrugs. 'Your front's pulled out. We'll renegotiate—or we walk.'''  

  \item[\textbf{Religious Order}] \index{religious order}\index{temple}  
        \emph{Chapters, rites, and influence.}  
        \textbf{GM Free Use Guidance:} Among believers, start \textbf{Dominant} or mitigate 1 SB from \emph{heresy/doubt}. \emph{Only} grant if the order is maintained and active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Call a council; force a pause in hostile religious proceedings. Describe the call: ``You strike the bell—'The Synod is convened. All chant.'''  
        \textbf{GM Neglect Hook:} If neglected, chapters secede or rites are corrupted. After 1 session neglected, start Dominant only every other session; after 2 sessions, it's Strained (requires 1 XP or a scene to restore order/rebuild trust).  
        \textit{Example:} ``The prioress signs the letter. 'The schism's widening. We'll hold the line—but don't expect unity.'''  
\end{description}

\section{Specialized Assets: GM Tools for Niche Power}
\label{sec:specialized-assets}
\index{assets!specialized}\index{GM!specialized assets}
Specialized assets are narrow tools with sharp edges. \textbf{As GM, use them to:}  
\begin{itemize}
  \item Reward specific, high-skill accomplishments with targeted leverage (e.g., a magical lab, an intelligence bureau).  
  \item Create hooks that tie into arcane, espionage, or technical complications.  
  \item Introduce complications that reflect the asset's fragility (e.g., lab accidents, blown covers).  
\end{itemize}
Each entry below includes \textbf{GM-specific guidance} on introduction, adjudication, and neglect consequences.

\begin{description}
  \item[\textbf{Magical Laboratory} (10 XP)] \index{laboratory!magical}\index{rituals}  
        \emph{Wards, benches, focus crystals.}  
        \textbf{GM Free Use Guidance:} Once per session, reduce Ritual Backlash by 1 SB (\S\ref{ch:magic}). \emph{Only} grant if the lab is maintained and wards are intact (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Contain an emergent effect for one beat; begin the control roll \textbf{Dominant}. Describe the containment: ``You slam the ward shut—'Hold!' The energy flickers like trapped lightning behind the sigils.''  
        \textbf{GM Neglect Hook:} If neglected, wards fade or equipment degrades. After 1 session neglected, reduce backlash only every other session; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to renew wards/replace gear).  
        \textit{Example:} ``The sigils are smudged. The crystal's cracked—and the air smells of ozone. Time for new wards—and a stern look at the apprentice.''  

  \item[\textbf{Training Academy} (12 XP)] \index{academy}\index{training}  
        \emph{Masters and mats.}  
        \textbf{GM Free Use Guidance:} In downtime, reduce \emph{one} skill advancement cost by 1 XP (min 1) for a designated trainee. \emph{Only} grant if the academy is maintained and trainee is progressing (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Call in a drilled team to assist off-screen; fill 1 segment on a \emph{Preparation} timer. Describe the arrival: ``Boots hit the dirt in unison—'Ready!' The trainees snap to attention.''  
        \textbf{GM Neglect Hook:} If neglected, masters leave or mats tear. After 1 session neglected, reduce skill cost only every other downtime; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to renew masters/replace mats).  
        \textit{Example:} ``The mat's split in two. One more fall—and it's useless. Time for new gear—and a lecture on safety.''  

  \item[\textbf{Intelligence Bureau} (10 XP)] \index{intelligence}\index{analysis}  
        \emph{Analysts and cut-outs.}  
        \textbf{GM Free Use Guidance:} Avoid surprise; the first recon opener starts \textbf{Dominant}. \emph{Only} grant if the bureau is maintained and analysts are active (track via downtime).  
        \textbf{GM Scene Surge Guidance:} Burn a channel; mitigate 1 SB from \emph{counterintelligence}. Describe the burn: ``You touch the wire—it sputters and dies. The channel's gone quiet.''  
        \textbf{GM Neglect Hook:} If neglected, analysts depart or channels fail. After 1 session neglected, avoid surprise only every other session; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to renew analysts/repair gear).  
        \textit{Example:} ``The tape's snarled. The signal's weak—and the cut-outs are missing three names. Time for new wire—and a headache.''  
\end{description}

\section{Tier III Institutions (16--20 XP): GM Tools for Empire-Shaping Power}
\label{sec:tier-iii-assets}
\index{assets!tier iii}\index{GM!tier iii assets}
Tier III institutions are empire-shaping engines requiring ongoing political play. \textbf{As GM, use them to:}  
\begin{itemize}
  \item Reward pinnacle accomplishments with world-shaping influence (e.g., a grand forge-citadel, a great library).  
  \item Create hooks that tie into imperial politics, economics, or major faction wars.  
  \item Introduce complications that scale with neglect (e.g., invasion, famine, heresy).  
\end{itemize}
Each entry below includes \textbf{GM-specific guidance} on introduction, adjudication, and neglect consequences.

\begin{description}
  \item[\textbf{Grand Forge-Citadel} (18 XP)] \index{forge}\index{citadel}  
        \emph{City heat and hammer.}  
        \textbf{GM Free Use Guidance:} Treat a batch of masterwork gear as immune to the first \emph{gear} complication this arc. \emph{Only} grant if the citadel is maintained and operational (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Rush a critical piece; fill 2 segments on a \emph{Build/Arm} timer. Describe the rush: ``The forge blazes as you hammer—sparks flying like angry stars. By dawn, it's done.''  
        \textbf{GM Neglect Hook:} If neglected, the citadel decays or workers revolt. After 1 session neglected, treat masterwork gear as immune only every other arc; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to renew infrastructure/appease workers).  
        \textit{Example:} ``The bellows are stiff. The hammer's head is loose—and the apprentices are whispering about walking out.'''  

  \item[\textbf{Great Library} (16 XP)] \index{great library}  
        \emph{Memory palace of the region.}  
        \textbf{GM Free Use Guidance:} Major inquiry begins \textbf{Dominant} and gains a \emph{Boon}. \emph{Only} grant if the library is maintained and the inquiry is valid (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Produce a sealed citation; negate a single \emph{misinterpretation} complication. Describe the citation: ``You break the seal—'Here, in the Master's own hand.' The words glow in the candlelight.''  
        \textbf{GM Neglect Hook:} If neglected, the library decays or scholars depart. After 1 session neglected, major inquiry begins at Controlled; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to restore infrastructure/replace scholars).  
        \textit{Example:} ``The shelves are sagging. The ink's faded—and the rats are nesting in the stacks.'''  

  \item[\textbf{Grand Exchange} (20 XP)] \index{exchange}\index{bourse}  
        \emph{Markets move when you breathe.}  
        \textbf{GM Free Use Guidance:} In high finance, ignore the first \emph{market shock}. \emph{Only} grant if the exchange is maintained and active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Halt trading bells for one exchange; DV $-1$ on one decisive deal. Describe the halt: ``You slam the gavel—'Trading is suspended. Effective immediately.' The floor falls silent.''  
        \textbf{GM Neglect Hook:} If neglected, the exchange falters or is challenged. After 1 session neglected, ignore market shock only every other arc; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to renew infrastructure/defend against challenges).  
        \textit{Example:} ``The clerk checks the ledger. 'The market's volatile today. Pay the premium—or we wait for calm.'''  

  \item[\textbf{Fleet Command} (18 XP)] \index{admiralty}\index{naval command}  
        \emph{Signals across sea lanes.}  
        \textbf{GM Free Use Guidance:} Declare blockade/corridor; hostile egress gains $+1$ step of difficulty. \emph{Only} grant if the command is maintained and signals are honored (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Redirect a squadron; improve Position for a maritime raid or escape. Describe the redirect: ``You signal the flagship—'New course bearing 270.' The fleet turns as one, prows cutting the waves.''  
        \textbf{GM Neglect Hook:} If neglected, the command decays or signals lapse. After 1 session neglected, declare blockade only every other session; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to renew infrastructure/defend signals).  
        \textit{Example:} ``The signalman eyes the broken mast. 'We can't send flags until it's fixed. Pay for new spars—or we drift.'''  

  \item[\textbf{Hydraulic Works} (16 XP)] \index{aqueduct}\index{irrigation}  
        \emph{Water is wealth.}  
        \textbf{GM Free Use Guidance:} Cancel the first \emph{shortage/drought} fallout this arc. \emph{Only} grant if the works are maintained and flowing (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Flood or feed a field at the critical moment; fill 1 segment on a \emph{Famine Relief} timer. Describe the flood: ``You open the gate—water rushes forth, sweet and cold. The field drinks deep.''  
        \textbf{GM Neglect Hook:} If neglected, the works clog or flow dwindles. After 1 session neglected, cancel shortage/drought only every other arc; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to clear clogs/restore flow).  
        \textit{Example:} ``The foreman points to the crack. 'One more freeze—and we're leaking. Pay for new stone—or we watch the crops die.'''  

  \item[\textbf{Palace Complex} (20 XP)] \index{palace}\index{court}  
        \emph{Keys to office and omen.}  
        \textbf{GM Free Use Guidance:} Political openers start \textbf{Dominant}. \emph{Only} grant if the palace is maintained and the court is active (track via downtime/Obligation).  
        \textbf{GM Scene Surge Guidance:} Convene a chamber; convert one \emph{scandal} consequence into a public inquiry you can contest. Describe the convening: ``You strike the gavel—'The Chamber is in session. All rise.'''  
        \textbf{GM Neglect Hook:} If neglected, the palace decays or court factions secede. After 1 session neglected, political openers start at Controlled; after 2 sessions neglected, it's Strained (requires 1 XP or a scene to renew infrastructure/mediate factions).  
        \textit{Example:} ``The gardener sighs at the wilted roses. 'One more neglect—and we're digging graves instead of growing joy.'''  
\end{description}

\section{Using Assets in Play: GM Procedures}
\label{sec:using-assets}
\index{assets!in play}\index{downtime}\index{flashbacks}
\textbf{As GM, your role is to:}  
\begin{itemize}
  \item Frame asset benefits \emph{only when the fiction supports it} (e.g., a safehouse is only useful if prepared and stocked).  
  \item Allow flashbacks \emph{only with a modest cost} (1 Boon or 1 downtime scene) to show the asset's prior work—\emph{never} as free exposition.  
  \item Tie complications \textbf{directly to asset condition}: If neglected, spend SB against it first; if compromised, deny benefits until fixed.  
  \item \textbf{Spotlight spread}: Rotate which assets take the stage so each investment breathes on-screen now and then—\emph{never} let one asset dominate all scenes.  
\end{itemize}

\paragraph{Example of Play (GM Perspective)}
Rellan uses his \textbf{Harbor Office (Standard Asset)}. \textbf{As GM:}  
\begin{itemize}
  \item \textbf{Frame the benefit:} At scene start, ask: ``Is the harbor office maintained and honored?'' If yes, grant free use to start customs inspection at \textbf{Dominant} position. Describe: ``The harbormaster nods at your seal—'All clear. Proceed to Berth 7.'''  
  \item \textbf{Allow flashback:} If Rellan wants to flash back to preparing the office, require 1 Boon or 1 downtime scene (e.g., ``You spent yesterday greasing the chains and bribing the night watch.'').  
  \item \textbf{Scene Surge:} When a rival tries to trap him on the wharf, Rellan spends a Boon for a Scene Surge. \textbf{As GM:} Describe the intervention vividly: ``You grip the chain lever—'Hold!'' The great iron links groan into place, blocking the channel. For one beat, the pursuit by water is negated—just enough to slip into a waiting skiff.''  
  \item \textbf{Neglect consequence:} If Rellan neglects the harbor office for 2+ sessions, advance a timer (e.g., \emph{Harbor Silt [4]} ticks +1 per session neglected) or spend 1--2 SB to create complications (e.g., ``The harbormaster frowns: 'Berth 3's silted up. Pay for dredging—or we wait for spring.''').  
\end{itemize}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Asset Quick Reference (GM),fonttitle=\bfseries,breakable]
\begin{itemize}
  \item \textbf{Costs}: Minor 4 XP / Standard 8 XP / Major 12 XP / Tier III 16--20 XP.  
        \textit{GM Guidance:} Adjust slightly for narrative weight (e.g., a unique artifact might cost 6 XP as a Minor Asset).  
  \item \textbf{Free Use}: Once per session, off-screen leverage \emph{only if fiction supports it}. Frame at scene start or as flashback with GM assent (cost: 1 Boon or 1 downtime scene).  
  \item \textbf{Scene Surge}: Spend 1 Boon for a one-beat, on-screen intervention. \textbf{Limit to one meaningful use per scene}.  
  \item \textbf{Upkeep}: Pay XP = base cost \emph{OR} devote a downtime scene. Neglect degrades condition—\textbf{use this as your complication engine}.  
  \item \textbf{Capacity (Optional)}: Assets $\leq$ \textbf{Spirit} are easy; beyond that, upkeep taxes mount or requires extra downtime scenes.  
\end{itemize}
\end{tcolorbox}

\section{Summary}
\index{assets!summary}\index{GM!assets summary}
Minor assets grease doors; Standard ones bend routes; Major holdings set tables; Tier III institutions tilt the map. \textbf{As GM:}  
\begin{itemize}
  \item Introduce assets as \textbf{narrative rewards}, not automatic power-ups.  
  \item Price them to reflect \textbf{story weight}, not just XP.  
  \item Always attach a \textbf{hook} (what happens if tended? what if neglected?).  
  \item Use asset \textbf{neglect} to drive complications—\textbf{never} let it be "nothing happens."  
  \item \textbf{Spotlight spread}: Rotate asset focus so each investment feels earned and consequential.  
\end{itemize}
Choose what to own, then let the world answer—and remember: the best assets create as many problems as they solve.

\begin{tcolorbox}[colback=black!3!white,colframe=black!60!white,title=From Faction to Asset: Taking Over a Terrestrial Patron (GM Guidance),label=sec:takeover-sidebar,breakable]
\index{terrestrial patrons!takeover}\index{assets!from factions}\index{GM!takeover}
Sometimes you stop owing the guild -- and start owning it. Rising to lead a faction (Thieves' Guild, Mercenary Company, Banking Charter, etc.) is a major story arc, not a simple XP purchase. \textbf{As GM, run this arc as a 2--3 session collaboration with the player.}  

\paragraph{Narrative Prerequisites (GM Verification)}
Before approving a takeover attempt, verify:  
\begin{itemize}
  \item The player has at least \textbf{6 Obligation} with the faction (deep entanglement).  
  \item They completed a \textbf{major service} for the faction (your discretion--e.g., saving it from destruction, winning a war, exposing a traitor at great cost).  
  \item They secured at least one \textbf{internal ally} with significant influence (another captain, a guildmaster, a wealthy backer).  
\end{itemize}
\textit{If prerequisites aren't met, the takeover attempt fails narratively--no XP cost, just a complication.}

\paragraph{The Takeover Arc (GM Collaboration)}
Collaborate on a 2--3 session arc that typically involves:  
\begin{enumerate}
  \item \textbf{Crisis or Power Vacuum} -- The current leader dies, disappears, or is disgraced (you define the crisis).  
  \item \textbf{Claim of Right} -- The player presents a valid claim (charter, election, duel victory, popular support--you judge validity).  
  \item \textbf{Consolidation} -- Defeat or co‑opt rivals, bribe or intimidate dissenters, and secure the faction's resources (you play the opposition).  
\end{itemize}
\textit{This arc should test the player's ingenuity, not just their XP pool.}

\paragraph{Mechanical Cost (GM Enforcement)}
Once the arc is complete, the player must spend \textbf{12 XP} (the cost of a Major Asset) to convert the faction into a personal Asset. \textbf{As GM:}  
\begin{itemize}
  \item Enforce this cost strictly--no conversions without payment.  
  \item Describe the conversion as institutional restructuring: ``The ledgers are rewritten, the oaths renewed--it's yours now, but the work has only begun.''  
\end{itemize}

\paragraph{Converting the Asset (GM Guidance)}
The new Asset uses the statistics of the most appropriate existing Asset from this compendium, modified by the faction's nature:  
\begin{itemize}
  \item \textbf{Thieves' Guild} $\rightarrow$ \textbf{Informant Ring} (Standard) or \textbf{Smuggler's Network} (Standard)  
  \item \textbf{Mercenary Company} $\rightarrow$ \textbf{Mercenary Contract} (Minor/Standard) or \textbf{Military Outpost} (Major)  
  \item \textbf{Banking Charter} $\rightarrow$ \textbf{Banking Charter} (Major) or \textbf{Trading Consortium} (Major)  
  \item \textbf{Royalist Network} $\rightarrow$ \textbf{Intelligence Bureau} (Specialized) or \textbf{Diplomatic Enclave} (Major)  
  \item \textbf{Ecktorian Mystery Cult} $\rightarrow$ \textbf{Temple Complex} (Standard) or \textbf{Magical Laboratory} (Specialized)  
\end{itemize}
\textit{As GM, feel free to create a custom Asset using \S\ref{sec:creating-your-own-asset} if none fit the faction's nature.}

\paragraph{What You Gain (GM Description)}
Upon conversion, grant:  
\begin{itemize}
  \item The Asset's normal \textbf{Free Use} and \textbf{Scene Surge} benefits.  
  \item The \textbf{Access Tag} associated with the faction (e.g., \textit{Guild-Ledgered}, \textit{Harbor Handsigns}).  
  \item A permanent \textbf{obligation track} toward the faction's members (now your underlings). This track is smaller (4 segments) and fills when you neglect their welfare or make reckless decisions that endanger them.  
\end{itemize}
\textit{As GM, track this obligation and use it to create complications when neglected (e.g., 'The crew grumbles--pay is late again. After 2 sessions neglected, they threaten to walk.')}

\paragraph{What You Lose (GM Enforcement)}
Enforce these losses strictly:  
\begin{itemize}
  \item The player can no longer use the faction as a \emph{terrestrial patron} to gain Obligation‑free perks--they are now responsible for it.  
  \item The faction becomes a target for your enemies--advance a \textbf{Rival Faction timer} whenever they use the Asset aggressively.  
  \item They must pay normal \textbf{upkeep} (XP or downtime scenes) to keep the Asset \emph{Maintained}. Neglect causes internal dissent, corruption, or defections--\textbf{use this as your complication engine}.  
\end{itemize}

\paragraph{Optional: Gradual Takeover (GM Permission)}
For a slower, more story‑driven approach, the player may forgo the 12 XP cost but must instead \textbf{spend a downtime scene per session} managing the faction (equivalent to Intensive upkeep) and accept that the Asset's benefits are unreliable until they have successfully completed a full arc of consolidation. \textbf{As GM, you may allow the XP cost to be paid in installments (e.g., 4 XP per downtime period) to represent gradual investment.}  

\paragraph{Example: Rellan Takes the Black Ledger (GM Narration)}
Rellan has served the Black Ledger smuggling syndicate for two arcs, accruing 6 Obligation. After exposing a rival who betrayed the syndicate (your narration), the aging leader steps down. Rellan spends 12 XP, completes a short arc consolidating loyal captains (you play the opposition), and converts the syndicate into a \textbf{Smuggler's Network (Standard Asset)}. He now commands hidden routes and safehouses--but must keep the crew paid and protected, or an internal revolt timer will start ticking.  
\textit{As GM, describe the takeover's cost: 'The syndicates' marks are scratched from the walls. New oaths are sworn--and new enemies are made.'}  
\end{tcolorbox}
