\chapter{Campaigns, Timers, and Consequences}
\label{ch:campaigns-timers}
\index{campaigns}\index{timers}\index{consequences}
\index{travel!GM guidance}

In \textbf{Fate's Edge}, campaigns are not a sequence of disconnected adventures---they are \textbf{pressure systems}.  
Every choice the players make pushes on something: a faction, a threat, a relationship, a belief. Over time, those pressures accumulate, shift, and eventually \emph{break}.

Your job as GM is not to hide this pressure---it is to \textbf{shape it, reveal it, and let the players decide where it lands}.

This chapter presents the core tools for doing that work: \textbf{Timers}, \textbf{Fronts}, \textbf{the Crown Spread}, and long-term campaign structures. Used well, they create momentum without railroading and stakes without arbitrary escalation.

\section{Foundational Principles for Campaign Design}

Before diving into tools, internalize these rules of thumb. They will save you more prep time than any other advice.

\subsection{Set Situations, Not Plots}
\index{campaigns!situations vs plots}

A \textbf{situation} is a stable condition with inherent conflict. A \textbf{plot} is a sequence of events you expect the players to follow.

\begin{itemize}
  \item \textbf{Plot:} ``The players must retrieve the Sunstone from the Ruined Tower, then defeat the necromancer, then close the rift.''
  \item \textbf{Situation:} ``The necromancer has claimed the Ruined Tower. The Sunstone inside could seal the rift. The village is starving because crops wither near the rift.''
\end{itemize}

Situations give players \textbf{agency}. They can negotiate with the necromancer, steal the Sunstone, collapse the tower, divert a river to flood the rift---any solution that fits the fiction.

\subsection{Design Timers, Not Every Element}
\index{timers!design principle}

You do not need to know exactly how a faction will act in session 12. Instead, give them a \textbf{timer} with a trigger condition.

\begin{quote}
\emph{``The Cult of the Eternal Night [6] advances when the players fail to protect a solar shrine or when the GM spends 2+ SB on occult activity.''}
\end{quote}

When the timer fills, the cult achieves its goal---and you improvise the specifics based on the current fiction.

\subsection{Read Your Player Sheets}
\index{player sheets!as campaign fuel}

Your players have already given you the best campaign hooks. Look for:
\begin{itemize}
  \item \textbf{Bonds:} ``I owe my life to the Vault-Queen's wardens.'' $\rightarrow$ The Vault-Queen calls in the debt.
  \item \textbf{Complications:} ``A sealed hold refuses you entry.'' $\rightarrow$ That hold is now a key location.
  \item \textbf{Patrons:} Any Patron choice is a promise to involve that Patron's agenda.
  \item \textbf{Skills at 3+}: If a player invested in Lore 3, they want to discover ancient secrets.
\end{itemize}

\subsection{Use Obligation as a Story Engine}
\index{Obligation!as story fuel}

Obligation (to Patrons, terrestrial patrons, or even the party's own Trust) is not a punishment---it is a \textbf{guarantee of future adventure}. When a player marks Obligation, note it. When their capacity is near full, introduce a demand.

\begin{quote}
\emph{``Your Patron, the Traveler, appears in a dream. The road to Highpass is washed out, and a pilgrim family is trapped. Guide them safely, and clear 2 Obligation. Refuse, and mark Fatigue 2---the road will not open for you again until you do.''}
\end{quote}

\subsection{Embrace the Cascade}
\index{consequences!cascade}

A single failure should never end the campaign. Instead, let it \textbf{cascade} into new problems.

\begin{quote}
\textit{The party fails to stop the ritual. The demon is summoned. Now the campaign is not ``stop the ritual'' but ``contain the demon before it corrupts the capital.''}
\end{quote}

Each failure should change the situation, not end it.

\subsection{Let Timers Be Visible}
\index{timers!visibility}

Timers are not hidden GM tools. Show them to the players. Write \emph{Cult Influence [6]} on an index card and place it on the table. When you tick it, announce: ``The cult's ritual advances. You feel the ground tremble.''

Visible timers create \textbf{tension} and \textbf{informed choices}. Players can prioritize which timer to slow.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Physical Trackers for Timers]
\index{trackers!physical}
Use tactile tools to make timers visible and engaging:
\begin{itemize}
  \item \textbf{Dry-erase boards or laminated sheets} with pre-drawn timer circles (4/6/8/10 segments). Mark ticks with a wet-erase marker.
  \item \textbf{Index cards with pie-chart wedges} drawn in pen; use a paperclip to mark current segment.
  \item \textbf{Poker chips or glass beads} placed on a numbered track from 0 to timer size.
  \item \textbf{Magnetic whiteboard with a timer template} and magnetic tokens.
\end{itemize}
Place the tracker in the center of the table where all players can see it. Tick it aloud: ``\emph{Ritual Completion} now at 3 of 6 segments.''
\end{tcolorbox}

\subsection{Ask Questions, Then Use the Answers}
\index{player input}

During Session Zero and between sessions, ask:
\begin{itemize}
  \item ``What does your character fear will happen if you fail this mission?''
  \item ``Who in your past might try to stop you?''
  \item ``What secret are you hiding from the rest of the party?''
\end{itemize}
Use these answers to seed complications. The player who fears losing their family? Their family is now in danger.

\subsection{Design for Your Tier}
\index{tiers!campaign design}

\begin{itemize}
  \item \textbf{Tier I--II (0--90 XP):} Local scope, 3--5 session arcs. Threats affect a village, a guild, a noble house.
  \item \textbf{Tier III (91--150 XP):} Regional scope, 6--10 session arcs. Threats affect a city, a province, a trade route.
  \item \textbf{Tiers IV--V (151+ XP):} Continental scope, 15+ session arcs. Threats affect nations, the natural order, reality itself.
\end{itemize}
Do not give a Tier I party a world-ending prophecy. Give them a rival who stole their map.

\section{Core GM Mechanics}

Understanding the core resolution loop is essential for effective campaign design. Every meaningful action follows this cycle:

\begin{enumerate}
  \item Declare intent and approach (Attribute+Skill)
  \item GM sets DV (2--5+) and Position (Dominant/Controlled/Desperate)
  \item Roll dice pool (Attribute+Skill d10s)
  \item Count successes (6--9=1 success; 10=2 successes)
  \item Apply outcome using the Outcome Matrix
  \item GM spends Story Beats (SB) to introduce complications
\end{enumerate}

\subsection{Difficulty Values (DV)}
\begin{itemize}
  \item \textbf{DV 2:} Routine--clear path, no pressure
  \item \textbf{DV 3:} Pressured--mild opposition (default)
  \item \textbf{DV 4:} Hard--serious resistance
  \item \textbf{DV 5+:} Extreme--dramatic gamble
\end{itemize}
Default DV is 3. Don't inflate DV to ``challenge'' players--use Position instead.

\subsection{Position}
Position tells you how risky the action is and changes how dice behave:
\begin{itemize}
  \item \textbf{Dominant:} You press advantage--re-roll one failure
  \item \textbf{Controlled:} Balanced normal risk--no re-rolls
  \item \textbf{Desperate:} You act under duress--re-roll one success
\end{itemize}
Spending 1 Boon improves Position by one step before rolling. The GM may spend SB to worsen Position.

\subsection{Outcome Matrix}
\begin{itemize}
  \item \textbf{Successes $\geq$ DV, SB=0:} Clean Success--Intent achieved, no complication
  \item \textbf{Successes $\geq$ DV, SB>$>$0:} Success with SB--Intent achieved; GM spends SB
  \item \textbf{0 $<$ Successes $<$ DV:} Partial--Progress; gain 1 Boon
  \item \textbf{Successes = 0:} Miss--No progress; gain 2 Boons; GM spends SB
\end{itemize}
A Partial is not failure--it moves the story forward. A Miss is never ``nothing happens.'' The GM must introduce a complication.

\subsection{Resources: Boons and Story Beats}
\subsubsection{Boons (Player Resource)}
\begin{itemize}
  \item Earned on Partial (1 Boon) or Miss (2 Boons)
  \item Maximum 5 Boons held; reduce to 2 at scene end (excess lost)
  \item Spent to: re-roll a die, activate an Asset, or improve Position by one step
  \item During downtime: convert 2 Boons $\rightarrow$ 1 XP (max 2 XP/session)
\end{itemize}

\subsubsection{Story Beats (GM Resource)}
\begin{itemize}
  \item Earned when players roll a 1 on any die
  \item Spent to introduce complications:
  \begin{itemize}
    \item 1 SB: Minor--noise, distraction, tick a 4--6 segment timer +1
    \item 2 SB: Moderate--alarm raised, lose advantage, worsen Position
    \item 3+ SB: Major--reinforcements, scene shifts, break an Asset
  \end{itemize}
  \item Be aggressive with SB--unused pressure is wasted momentum
\end{itemize}

\section{Campaign Tools}

\subsection{The Crown Spread (Session Zero)}
\index{Crown Spread}\index{campaign!setup}

The Crown Spread is a five-card (or six-card) oracular tool for seeding a campaign's major elements. Use it during Session Zero or at the start of a new arc. It creates a web of relationships, not a railroad.

\subsubsection{What You Need}
\begin{itemize}
  \item A standard 52-card deck (jokers optional).
  \item A table or whiteboard to record the spread.
  \item 15--20 minutes of quiet thinking.
\end{itemize}

\subsubsection{The Spread Positions}
Draw one card for each position. Lay them in a semicircle (the ``Crown'').

\begin{center}
\begin{longtable}{lll}
\caption{Spread Positions}\\
\toprule
\textbf{Position} & \textbf{Card Drawn} & \textbf{What It Represents} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{Position} & \textbf{Card Drawn} & \textbf{What It Represents} \\
\midrule
\endhead
\bottomrule
\endfoot
Spade & (e.g., 5\spadesuit) & \textbf{The Crown Site} -- where destiny is decided. \\
Heart & (e.g., Q\heartsuit) & \textbf{The Crown Rival} -- who stands between the party and their goals. \\
Club & (e.g., 9\clubsuit) & \textbf{The Crown Pressure} -- the relentless force that prevents complacency. \\
Diamond & (e.g., A\diamondsuit) & \textbf{The Crown Leverage} -- the advantage that can turn the tide. \\
Wild (draw a 6th card) & (e.g., K\spadesuit) & \textbf{The Hidden Force} -- the unexpected element. \\
\end{longtable}
\end{center}

\subsubsection{Interpreting the Cards}
Use the suit to determine the nature of the element, then interpret the rank to refine it.

\begin{tcolorbox}[colback=gray!5!white, colframe=black, title=Quick Interpretation Guide]
\begin{description}
\item[Spade (Site)] -- A place: ruin, bridge, forest, tower, cave, city district, crossroads.
\item[Heart (Rival)] -- A person with emotional ties: family, former ally, rival, spurned lover, grieving widow.
\item[Club (Pressure)] -- A resource or systemic threat: famine, plague, debt, war tax, blight, curse.
\item[Diamond (Leverage)] -- An object or secret: a lost letter, a relic, a true name, a hidden passage, a witness.
\item[Wild (Hidden)] -- Anything: a monster, a traitor, a natural disaster, a forgotten god, a second faction.
\end{description}

\medskip
\noindent \textbf{Rank guidance:}
\begin{itemize}
  \item \textbf{Ace -- 3:} Minor scale, personal stakes.
  \item \textbf{4 -- 6:} Moderate scale, affects a community.
  \item \textbf{7 -- 10:} Major scale, threatens a region.
  \item \textbf{Jack -- King:} Face cards suggest a named individual with personality and agenda.
\end{itemize}
\end{tcolorbox}

\subsubsection{Connecting the Elements}
After the draw, ask yourself:
\begin{enumerate}
  \item How does the Rival relate to the Site?
  \item How does the Pressure affect the Site or Rival?
  \item How can the Leverage help against the Pressure or Rival?
  \item What does the Hidden Force want, and how does it complicate the others?
\end{enumerate}
Draw a simple relationship map. You now have a campaign skeleton.

\subsubsection{Timers for the Crown Spread}
Create one visible timer for each element that will take time to resolve. Use these defaults:

\begin{center}
\begin{longtable}{ll}
\caption{Sample Timers for Crown Elements}\\
\toprule
\textbf{Element} & \textbf{Sample Timer } \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{Element} & \textbf{Sample Timer } \\
\midrule
\endhead
\bottomrule
\endfoot
Rival's Agenda & \textit{[Rival Name]'s Consolidation} [6] \\
Pressure & \textit{[Pressure Name] Spreads} [8] \\
Leverage (if it can be lost) & \textit{[Leverage Name]'s Safety} [6] \\
Hidden Force (if it escalates) & \textit{[Hidden Name]'s Awakening} [4] \\
\end{longtable}
\end{center}

Tick these timers during play when players fail relevant rolls or ignore the element. When a timer fills, that element's situation worsens dramatically.

\subsubsection{GM Quick Cues}
\begin{itemize}
  \item \textbf{Don't over-plan.} The spread is a seed, not a script. Let players discover connections through play.
  \item \textbf{Be flexible.} If a better idea emerges, replace an element -- the cards are oracles, not chains.
  \item \textbf{Use the Wild card for surprise.} Even you don't have to know what it is until it appears.
  \item \textbf{Tie every character to at least one element.} During Session Zero, ask each player: ``How do you know about the Rival?'' or ``Why does the Pressure affect you personally?''
\end{itemize}

\subsection{Timers: The GM's Primary Tool}
\index{timers!GM guidance}

Timers are the single most useful tool in Fate's Edge. Use them for:
\begin{itemize}
  \item \textbf{Threats:} \emph{Reinforcements Arrive [4]}, \emph{Ritual Completes [6]}
  \item \textbf{Opportunities:} \emph{Alliance Forming [6]}, \emph{Weakness Discovered [4]}
  \item \textbf{Travel:} \emph{Journey to the Pass [8]}
  \item \textbf{Social:} \emph{The Duke's Patience [4]}, \emph{Rumor Spreads [6]}
\end{itemize}

\subsubsection{When to Tick a Timer}
\begin{itemize}
  \item On a \textbf{Miss} or \textbf{Partial} -- even if the player otherwise succeeds.
  \item When the GM spends \textbf{Story Beats} (1 SB = 1 tick typically).
  \item When the fiction demands it (e.g., ``While you argue, the fire spreads'').
  \item At the end of a session (for long-term timers like \emph{Cult Influence}).
\end{itemize}

\subsubsection{Timer Size Guidelines}
\begin{itemize}
  \item \textbf{4 segments:} Short-term tension (a patrol arrives, a door breaks).
  \item \textbf{6 segments:} Standard scene timer (a ritual completes, a rival escapes).
  \item \textbf{8 segments:} Multi-scene or session timer (a faction consolidates power).
  \item \textbf{10+ segments:} Campaign timer (a war, a prophecy, a curse).
\end{itemize}

\subsection{Factions and Fronts}
\index{factions}\index{fronts}

A \textbf{faction} is any organized group with goals that intersect the party's. A \textbf{front} is a collection of related factions and timers that represent a major threat.

\subsubsection{Faction Sheet (Quick Template)}

\begin{tcolorbox}[colback=white!97!gray, colframe=black!80!gray, title=Faction Template]
\begin{tabular}{p{4cm}p{10cm}}
\textbf{Name:} & \hrulefill \\
\textbf{Agenda Timer :} & [\hspace{1cm}] (what they want to achieve) \\
\textbf{Key NPCs:} & (2--3 names and roles) \\
\textbf{Resources:} & (what they can bring to bear) \\
\textbf{Ticking conditions:} & (when does their timer advance?) \\
\textbf{Visible tracks:} & (list timers players can see) \\
\end{tabular}
\end{tcolorbox}

\subsubsection{Running Faction Turns}
\index{factions!turn}

Between sessions (or at the start of a session), run a quick \textbf{faction turn} to keep the world alive. A faction turn takes 5--10 minutes.

\begin{enumerate}
  \item \textbf{List active factions.} Keep no more than 3--5 that directly interact with the party.
  \item \textbf{For each faction, roll 1d6.} Use the table below.
  \item \textbf{Advance or retreat their Agenda Timer .}
  \item \textbf{Note one concrete change.} Write one sentence per faction.
  \item \textbf{Read the changes aloud} at the start of the next session.
\end{enumerate}

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Faction Progress Roll (d6)]
\begin{center}
\begin{longtable}{cll}
\caption{Faction Progress Roll}\\
\toprule
\textbf{Roll} & \textbf{Progress} & \textbf{Agenda Timer Change} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{Roll} & \textbf{Progress} & \textbf{Agenda Timer Change} \\
\midrule
\endhead
\bottomrule
\endfoot
1--2 & Setback & --1 segment (min 0) \\
3--4 & Stalled & No change \\
5--6 & Advance & +1 segment \\
\end{longtable}
\end{center}
\end{tcolorbox}

\subsubsection{Optional: Faction Standing}
For factions that are not pursuing a single agenda but merely have an attitude, track a simple \textbf{-3 to +3 scale}:
\begin{itemize}
  \item \textbf{-3:} Enemy -- actively works against the party.
  \item \textbf{0:} Neutral -- indifferent.
  \item \textbf{+3:} Ally -- will sacrifice for the party.
\end{itemize}
Shift standing by $\pm 1$ after major interactions. No dice needed -- just fiction.

\subsection{Rival Parties}
\index{rival parties}

A rival party is a faction of 3--5 NPCs with their own Patrons, assets, and a visible \textbf{Rival Agenda timer} (6--8 segments). Tick their timer when the PCs ignore them or the GM spends SB on their off-screen activities. When the timer fills, the rivals achieve a major milestone.

Reveal their progress through:
\begin{itemize}
  \item \textbf{Rumors:} ``The Ashen Cord was seen leaving the Sunstone vault.''
  \item \textbf{Found scenes:} A burned camp, a dead informant, a signature left behind.
  \item \textbf{Direct confrontation:} They show up during a critical moment.
\end{itemize}

For a full template and example (The Ashen Cord), see Appendix \ref{app:rival-generator}.

\section{Complications and Hazards}
\index{complications}\index{hazards}

Complications are the narrative, mechanical, and environmental consequences that arise from failure, pressure, or the GM spending Story Beats. Hazards are ongoing complications that require action to avoid or mitigate.

\subsection{Using Story Beats for Complications}
When spending SB, follow this menu:
\begin{itemize}
  \item \textbf{1 SB:} Minor pressure--noise, distraction, tick a relevant timer +1, worsen Position by one step
  \item \textbf{2 SB:} Moderate setback--alarm raised, lose cover/advantage, lesser foe appears, split the party briefly
  \item \textbf{3 SB:} Serious trouble--reinforcements arrive, key gear breaks (temporarily), major twist shifts the scene
  \item \textbf{4+ SB:} Scene-shaping turn--a trap springs, authority arrives, the environment changes dramatically, a Patron manifests an omen
\end{itemize}

\subsection{Hazard Timers}
Represent active hazards (fire, flooding, toxic gas) with Threat Timers:
\begin{itemize}
  \item Tick the timer on missed rolls, GM SB spends, or at scene intervals
  \item When the timer fills, the hazard triggers (e.g., collapse, wave of fire)
  \item Characters may attempt to slow or disable the hazard (e.g., Tinker to seal a leak, Magic to freeze spreading flames)
  \item Treat as a Progress Timer when players work against it
\end{itemize}

\subsection{Social Hazards}
\begin{itemize}
  \item \textbf{Rumors:} A timer that advances when the party fails social rolls; when full, a faction turns hostile
  \item \textbf{Scrutiny:} Worsens Position for stealth or deception
  \item \textbf{Reputation Damage:} Future social rolls start at worse Position
\end{itemize}

\subsection{Magical Hazards}
\begin{itemize}
  \item \textbf{Backlash Zone:} An area where free casting automatically generates +1 SB per spell
  \item \textbf{Ward/Seal:} Requires a roll to pass; failure triggers a consequence (harm, alarm, etc.)
  \item \textbf{Leaking Veil:} Spirits or fey occasionally manifest spontaneously (GM spends SB to introduce)
\end{itemize}

\clearpage
\section{Settlement Generator}
\label{sec:settlement-generator}
\index{settlement generator}\index{generator!settlement}

Use this generator to create a memorable settlement---village, town, or city---in any region of \textit{Fate's Edge}. Draw cards or roll dice to determine size, dominant faction, a pressing problem, a local resource, and a regional quirk. Then add hooks that tie the settlement to your campaign.

\subsection{Step 1: Settlement Size}
\label{sec:settlement-size}
Roll 1d6 or draw a \spadesuit\ (Spade) and consult the rank.

\begin{center}
\begin{longtable}{cll}
\caption{Settlement Size}\\
\toprule
\textbf{d6} & \textbf{Size} & \textbf{Population \quad Notable Features} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Size} & \textbf{Population \quad Notable Features} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Thorp & 20--80 -- a handful of families, one well, no walls. \\
2 & Hamlet & 81--400 -- a crossroads inn, a small shrine, a mill. \\
3 & Village & 401--900 -- a weekly market, a temple, a wooden palisade. \\
4 & Town & 901--2,000 -- stone walls, guild hall, a modest lord's keep. \\
5 & Large Town & 2,001--5,000 -- multiple gates, a mint, a trade charter. \\
6 & Small City & 5,001--12,000 -- wards, a cathedral, a permanent garrison. \\
\end{longtable}
\end{center}

\subsection{Step 2: Dominant Faction or Power}
\label{sec:settlement-faction}
Roll 1d10 or draw a \heartsuit\ (Heart). Adjust for region as needed.

\begin{center}
\begin{longtable}{cl}
\caption{Dominant Faction or Power}\\
\toprule
\textbf{d10} & \textbf{Faction} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d10} & \textbf{Faction} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Noble House (baron, count, local strongman) \\
2 & Merchant Guild (spice factors, silk merchants, salt syndicate) \\
3 & Temple / Cult (Everflame Synod, Veiled Sister shrine, fire-temple) \\
4 & Wizard / Scholar (archivist, college dean, hedge-witch coven) \\
5 & Mercenary Company (condotta captain, free company garrison) \\
6 & Crime Lord (smuggler king, fence, a ``reeve'' who takes bribes) \\
7 & Council of Elders (moot, well-keepers, clan council) \\
8 & Religious Minority (heretical sect, hidden shrine, outlawed order) \\
9 & Foreign Enclave (Sidhi lighthouse keepers, Aeler engineers, Tulkani traders) \\
10 & The Dead (ghost council, cairn that votes, barrow-wight) \\
\end{longtable}
\end{center}

\subsection{Step 3: Settlement Problem}
\label{sec:settlement-problem}
Roll 1d10 or draw a \clubsuit\ (Club).

\begin{center}
\begin{longtable}{clp{7cm}}
\caption{Settlement Problem}\\
\toprule
\textbf{d10} & \textbf{Problem} & \textbf{Twist} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d10} & \textbf{Problem} & \textbf{Twist} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Food Shortage & The granary was sabotaged; the last caravan never arrived. \\
2 & Monster Threat & Beast(s) from the wilds attack livestock or people. \\
3 & Feud & Two families or factions are on the edge of bloodshed. \\
4 & Taxation / Debt & A lord or factor demands payment the village cannot make. \\
5 & Sickness & A plague, a curse, or poisoned water spreads. \\
6 & Cult Activity & The Ninth's whisperers, a demon-worshipping cell, or a doomsday sect. \\
7 & Unquiet Dead & Ghosts, drowned, or cairn-wights walk at night. \\
8 & Trade Route Cut & A bridge is down, a pass is blocked, or pirates raid. \\
9 & Betrayal & Someone trusted sold secrets, stole goods, or framed an innocent. \\
10 & The Hollow's Attention & The ninth cup was poured; the Oath-Keeper has been seen. \\
\end{longtable}
\end{center}

\subsection{Step 4: Local Resource or Leverage}
\label{sec:settlement-resource}
Roll 1d10 or draw a \diamondsuit\ (Diamond).

\begin{center}
\begin{longtable}{cll}
\caption{Local Resource or Leverage}\\
\toprule
\textbf{d10} & \textbf{Resource} & \textbf{What it provides} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d10} & \textbf{Resource} & \textbf{What it provides} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Salt Pans & Salt, preservation, trade with inland towns. \\
2 & Iron Mine & Ore, weapons, engineering contracts. \\
3 & Timber & Shipbuilding, charcoal, construction. \\
4 & Spice / Coffee & High-value cash crop, foreign trade. \\
5 & Honey / Wax & Sweetener, candles, medicine. \\
6 & Ward-Salt (Mistlands) & Protection against wraiths; valuable to outsiders. \\
7 & Silk Weaving & Fine cloth, witness cords, diplomatic gifts. \\
8 & Brewing / Wine & Ale, mead, or wine; a social hub. \\
9 & Relic / Shrine & Pilgrimage, healing, religious authority. \\
10 & Old Records / Maps & Knowledge of ruins, lost passes, or hidden vaults. \\
\end{longtable}
\end{center}

\subsection{Step 5: Regional Quirk}
\label{sec:settlement-quirk}
A settlement's quirk gives it an unforgettable flavour. Use one of the two methods below.

\subsubsection{Method A: Draw from the Regional Generator}
For a settlement in a specific region, draw a single card from that region's deck (or use a standard deck and interpret by suit). The suit tells you the quirk's nature; the rank tells you its intensity.

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Suit} & \textbf{The quirk manifests as...} \\
\midrule
\spadesuit & A notable place or landmark (e.g., a peculiar bridge, a haunted well, a strange stone circle). \\
\heartsuit & A person or faction with an unusual custom, obsession, or secret. \\
\clubsuit & A recurring complication or local superstition (weather, beasts, spirits). \\
\diamondsuit & A surprising resource, trade good, or piece of forbidden knowledge. \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Rank guidance:}
\begin{itemize}
  \item \textbf{2--5:} Minor, easily overlooked (a strange sign, a whispered rumour).
  \item \textbf{6--10:} Noticeable, affects daily life (a locked tower, a weekly bell that rings nine times).
  \item \textbf{Jack--King:} Major, defines the settlement (a noble's curse, a bridge that tolls memories).
  \item \textbf{Ace:} The Hollow is close; the ninth taboo is woven into the settlement's bones.
\end{itemize}

\textbf{Example:} In Thepyrgos, a draw of \textbf{9\clubsuit} might be a \emph{Cistern Taint} that glows green, or a \emph{Synod Summons} posted on every door. In Ubral, a \textbf{Queen\heartsuit} could be a \emph{Lady of Tor} who rules from a cairn, or a \emph{Bride-Carrier} who walks between feuding clans.

\subsubsection{Method B: d6 Quirks by Major Region}
If you prefer a quick roll, use the tables below. Each table offers six quirks drawn from the region's established flavour.

\paragraph{Kahfagia (Creole Thalassocracy)}
\begin{center}
\begin{longtable}{cl}
\caption{Kahfagia Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & The harbour has a ninth buoy that only appears at low tide; sailors avoid it. \\
2 & Every contract is witnessed by a Mirror-Keeper; the mirrors never show the same face twice. \\
3 & A house of the Shadow Council stands empty; silk threads cover the doorframe. \\
4 & Lanterns change colour with the tide; no one explains the code to outsiders. \\
5 & The local well tastes of salt and iron; the drowned are said to drink from it. \\
6 & A fog-bell rings at odd hours; those who count the rings never reach nine. \\
\end{longtable}
\end{center}

\paragraph{Ecktoria (Marble \& Fire)}
\begin{center}
\begin{longtable}{cl}
\caption{Ecktoria Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & A statue in the forum turns its head when someone lies. \\
2 & The Everflame in the local basilica burns blue on market days. \\
3 & Coin-house seals are stamped with a bell that rings when the debt is unpaid. \\
4 & Censor's clerks audit the dead as well as the living; names appear on rolls after burial. \\
5 & A Triumph Stair has a missing ninth step; petitioners who step into the gap are never seen again. \\
6 & The Amber Palace's quarantine is felt even here; patricians send messages by bell-code. \\
\end{longtable}
\end{center}

\paragraph{Acasia (Broken Marches)}
\begin{center}
\begin{longtable}{cl}
\caption{Acasia Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & A milestone points the wrong way; locals follow it anyway. \\
2 & The toll-bridge keeper is a child with no shadow; she never speaks. \\
3 & Hedge-witches tie red thread to every gate; the thread tightens when a promise is broken. \\
4 & Free company colours change each week; no one is sure who holds the contract. \\
5 & The Margravine's tally-stick is nailed to the inn wall; each notch is a paid debt. \\
6 & At dusk, the sound of legionary boots echoes from the old road. \\
\end{longtable}
\end{center}

\paragraph{Vhasia (Fractured Sun)}
\begin{center}
\begin{longtable}{cl}
\caption{Vhasia Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Every noble's seal has a hairline crack; no one remembers why. \\
2 & The local cantor sings the \emph{Planctus Armatus} each dusk; the ninth verse is silence. \\
3 & A mirror in the square shows not your face but the sack of Lence. \\
4 & Knights touch their gorgets before speaking of mercy; some unbuckle them at sunset. \\
5 & The river carries ash from the Belworth; farmers scatter it on their fields for luck. \\
6 & A child with no shadow offers directions; those who follow forget their names. \\
\end{longtable}
\end{center}

\paragraph{Viterra (Hedge-Law Realm)}
\begin{center}
\begin{longtable}{cl}
\caption{Viterra Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Every hedge has a tied ribbon; each colour represents a different law. \\
2 & The Queen's Justiciar visits once a season; she never draws her sword, only her seal. \\
3 & A border stone moves three feet each spring; surveyors argue over who moved it. \\
4 & Local courts settle disputes by hedge-walk; the party who gets lost first loses. \\
5 & The Fenwood comptroller's quill is a dagger; he uses it for both writing and reminders. \\
6 & A bell in the square rings eight times at dawn; the ninth is for the absent. \\
\end{longtable}
\end{center}

\paragraph{Thepyrgos (City of Unfinished Proofs)}
\begin{center}
\begin{longtable}{cl}
\caption{Thepyrgos Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & A stair that leads nowhere stands in the square; locals leave offerings. \\
2 & The town's bell rings eight times at noon, never nine. \\
3 & Every contract must be witnessed by a bell-ringer. \\
4 & A broken astrolabe hangs in the church, still ticking. \\
5 & The well water glows faintly green after rain. \\
6 & A Chain-Lantern patrol tests travellers at the gate; they ask for a name that rings false. \\
\end{longtable}
\end{center}

\paragraph{Aeler (Under-Vaults)}
\begin{center}
\begin{longtable}{cl}
\caption{Aeler Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Miners count their breaths aloud in eights; visitors who count nine are shunned. \\
2 & A sealed door in the oldest building is warm to the touch; no one opens it. \\
3 & The local smith's hammer strikes three times in the morning, three at night. \\
4 & Every household has a bell that rings when the mountain shifts. \\
5 & A breath-counter hangs in the inn; guests must register before drinking. \\
6 & The True Masons inspect the well each month; they never explain what they measure. \\
\end{longtable}
\end{center}

\paragraph{Ykrul (Wolf Standards)}
\begin{center}
\begin{longtable}{cl}
\caption{Ykrul Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Horses are never named; riders earn the right to name a foal after a raid. \\
2 & A hostage string hangs in the centre of camp; each knot is a kept promise. \\
3 & The bone-singer sings the names of the dead at dawn; the dead sing back. \\
4 & Salt is measured in handfuls, not coin; a closed fist means war. \\
5 & The Khagan's station has a ninth felt layer beneath the throne; no one has seen it. \\
6 & White squalls are greeted with silence; speaking during one invites the Hollow. \\
\end{longtable}
\end{center}

\paragraph{Vilikari (Laurels \& Longhouses)}
\begin{center}
\begin{longtable}{cl}
\caption{Vilikari Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & The local longhouse has a marble bench beside the oak; both are used for judgments. \\
2 & Every contract is signed with a braid of hair as well as a seal. \\
3 & A Foedus Stone fragment is kept in the square; its newest crack has no date. \\
4 & The innkeeper pours eight cups; the ninth is left for the witness no one wants. \\
5 & Border riders speak a pidgin of Latin and steppe cant; outsiders cannot follow. \\
6 & A Vilikari factor keeps two ledgers: one for Ecktoria, one for the clan. \\
\end{longtable}
\end{center}

\paragraph{Linns (Skerries \& Storm-Oaths)}
\begin{center}
\begin{longtable}{cl}
\caption{Linns Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Every boat has a ninth oar that is never used; it is for the drowned. \\
2 & A fog-bell rings at the harbour mouth; those who hear nine rings must stay ashore. \\
3 & The Thing-Holm ice never melts; oaths carved into it are frozen forever. \\
4 & Skalds sing verses in eights; the ninth line is always a question. \\
5 & A net-wife ties a knot for every storm she survives; her belt is thick with knots. \\
6 & The Sea-Queen's ship has a green lantern at the mast; it is never lit. \\
\end{longtable}
\end{center}

\paragraph{Ubral (Stone Between Spears)}
\begin{center}
\begin{longtable}{cl}
\caption{Ubral Quirks}\\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{Quirk} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & A large cairn stands at the edge of town; widows add a stone every year. \\
2 & The inn has a guest-cup that is never washed; it holds the memory of every traveller. \\
3 & Cattle are counted every morning; a missing calf starts a feud. \\
4 & A dwarf outcast runs the forge; he speaks only in eights. \\
5 & The village has no jail; criminals are tied to a stone outside for three days. \\
6 & A bean-nighe (washer ghost) appears at the ford before a death. \\
\end{longtable}
\end{center}

\subsection{Step 6: Hooks}
\label{sec:settlement-hooks}
Every settlement should offer immediate reasons to engage. Roll twice (or pick from each column).

\begin{center}
\begin{longtable}{cll}
\caption{Settlement Hooks}\\
\toprule
\textbf{d6} & \textbf{What the settlement needs (external)} & \textbf{What is hidden / secret} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{d6} & \textbf{What the settlement needs (external)} & \textbf{What is hidden / secret} \\
\midrule
\endhead
\bottomrule
\endfoot
1 & Protection from raiders / monsters & A traitor in the leader's inner circle. \\
2 & A lost heirloom or charter (stolen or taken) & A forbidden shrine to the Ninth or a dark power. \\
3 & A mediator for a feud (two parties at each other's throats) & A spy from a rival faction (Admiralty, Censorate, etc.) \\
4 & A new trade route or a bridge repaired & A hidden vault with old records or treasure. \\
5 & Delivery of a sealed message or a dangerous prisoner & A murder that implicates an innocent (or a ghost). \\
6 & A witness to a crime (someone saw something they shouldn't) & The Hollow's mark -- something is counting the settlement's days. \\
\end{longtable}
\end{center}

\subsection{Example: Rolling a Settlement}
\label{sec:settlement-example}

\textbf{Draw / Roll}
\begin{itemize}
  \item Size: 3 $\rightarrow$ \textbf{Village} (401--900)
  \item Faction: 9 $\rightarrow$ \textbf{Foreign Enclave} (Sidhi lighthouse keepers)
  \item Problem: 7 $\rightarrow$ \textbf{Unquiet Dead} (ghosts walk at night)
  \item Resource: 4 $\rightarrow$ \textbf{Spice / Coffee} (high-value trade)
  \item Regional Quirk (Thepyrgos, d6): 3 $\rightarrow$ \textbf{Every contract must be witnessed by a bell-ringer.}
  \item Hooks: 1 + 5 $\rightarrow$ \textbf{Protection from raiders} + \textbf{A hidden vault with old records}
\end{itemize}

\textbf{Settlement: Wharfstead of the Green Lamp} (Thepyrgos borderlands)  

\noindent A Sidhi enclave in a dry river valley, surrounded by olive terraces. The Sidhi keep a small lighthouse that burns green, witnessing every contract. At night, drowned ferrymen from the Black River walk the streets, asking for passage home. The local coffee is famous, but a band of reivers has been raiding the warehouses. In the lighthouse basement, a vault holds Astroegro navigation charts -- and a fragment of a forbidden record.

\subsection{One-Card Quick Settlement (For Improv)}
\label{sec:settlement-quick}

If you need a settlement \emph{right now}, draw one card.

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Suit} & \textbf{Element} \\
\midrule
\spadesuit & The place: a landmark or building (bridge, mill, shrine, tower). \\
\heartsuit & The power: a person or faction in charge. \\
\clubsuit & The trouble: what is broken, stolen, or threatening. \\
\diamondsuit & The resource: what the settlement produces or trades. \\
\bottomrule
\end{tabular}
\end{center}

\noindent \textbf{Rank} determines severity (low = local, high = regional). \textbf{Face card} means the settlement has a named leader with a secret. \textbf{Ace} means the Hollow is watching -- the ninth taboo has been broken recently.

\index{settlement generator!example}
\index{Wharfstead of the Green Lamp}

\clearpage

\section{Running Travel as a Medieval Trek}
\label{sec:gm-travel}
\index{travel!GM techniques}

Travel in Fate's Edge is not a line on a map. It is a \textbf{medieval journey} -- uncertain, weather-beaten, dependent on local guides, hospitality, and the kindness (or cruelty) of strangers. For full procedures (weather tables, road condition dice, regional decks), see the \textit{Travel Guide} and \textit{Player's Guide} Chapter 10. This section gives you the GM principles.

\subsection{Lead with Weather, Not Distance}
Open each leg with a weather description, then set Position based on that weather (e.g., Desperate for navigation in fog, Controlled for stealth in rain). The weather \emph{is} the terrain.

\subsection{Maps Are Lies; Guides Are Truth}
Require a \textbf{Guide} role for each leg (PC or hired NPC). The Guide's roll (Wits + Survival or Lore) determines if they find the safe route, the short route, or the one that leads to a cultist ambush.

\subsection{Hospitality as a Resource}
When the party seeks shelter, draw a \textbf{Heart} from the regional deck -- that's the host. Their attitude affects what they offer. Spending a \textbf{String} (e.g., ``I helped this village last year'') can improve the outcome.

\subsection{The Travel Timer as Mood Tracker}
Do not just tick the timer on successes. Use it to track the \emph{feeling} of the journey (Optimism $\rightarrow$ Fatigue $\rightarrow$ Desperation $\rightarrow$ The Edge). When the timer fills, describe the arrival -- mud-caked and grateful.

\subsection{Genre Shifts Through Borders}
When the party crosses from one region to another, the dominant fantasy subgenre shifts. Use the travel deck to introduce the new region's flavor immediately (e.g., from Ecktoria's political intrigue to the Mistlands' folk horror: the road grows silent, fog rolls in, a bell tolls without wind).

\subsection{Traveling Without a Map}
Fate's Edge does not use hex crawls. Travel is abstracted into \textbf{legs}, each with a Travel Timer (4, 6, 8, or 10 segments). You need only know how many \textbf{meaningful scenes} stand between two locations -- not the miles.

\subsection{Roles on the Road}
Assign one role per character each leg. Roles may rotate each leg.

\begin{center}
\begin{longtable}{lll}
\caption{Travel Roles and Benefits}\\
\toprule
\textbf{Role} & \textbf{Key Test} & \textbf{Benefit on Success} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{Role} & \textbf{Key Test} & \textbf{Benefit on Success} \\
\midrule
\endhead
\bottomrule
\endfoot
Guide & Wits+Survival & Tick the Travel Timer +1 \\
Scout & Wits+Stealth & Party cannot be surprised on the first encounter of the leg \\
Quartermaster & Wits+Craft & Prevent one Supply drop or recover 1 spent Supply \\
Watch & Presence+Notice & Party gains +1 Position on the first defense roll of any ambush \\
\end{longtable}
\end{center}

Each role action requires a roll (Attribute+appropriate skill, DV set by the GM). Failure may tick the Travel Timer backwards or introduce a complication.

\subsection{Encounters and Events}
Every 1--2 segments filled, the GM introduces an encounter or hazard. At minimum, each leg should include one meaningful encounter.

\begin{center}
\begin{longtable}{lll}
\caption{Encounter Types}\\
\toprule
\textbf{Type} & \textbf{Examples} & \textbf{Typical Resolution} \\
\midrule
\endfirsthead
\multicolumn{3}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{Type} & \textbf{Examples} & \textbf{Typical Resolution} \\
\midrule
\endhead
\bottomrule
\endfoot
Discovery & Ruins, strange omens, hidden sites & Investigation or exploration roll \\
Hazard & Terrain challenge, illness, extreme weather & Survival or Endurance test; may use a hazard timer \\
Encounter & Bandits, emissaries, beasts, rival travelers & Combat or social resolution \\
Complication & Bond tested, supplies strained, pursuit gains ground & Roleplay or skill challenge \\
\end{longtable}
\end{center}

On a failed travel roll, the GM may spend SB to introduce an encounter immediately.

\subsection{Supply and Depletion}
Track shared party supplies with a Supply Timer [4]:
\begin{itemize}
  \item Tick the timer when the Quartermaster fails a roll, when the GM spends SB on scarcity, or when the party is forced to march without resupply
  \item When the timer fills, the party runs out of a critical resource (food, water, fuel)
  \item Apply a Condition (Hungry, Thirsty, etc.) or mark 1 Fatigue per character
  \item Resupply at settlements, through successful foraging, or by spending a Cache token (see Asset rules)
\end{itemize}

\subsection{Weather and Terrain Modifiers}
The GM may apply modifiers based on conditions:

\begin{center}
\begin{longtable}{ll}
\caption{Weather and Terrain Modifiers}\\
\toprule
\textbf{Condition} & \textbf{Effect} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\emph{Continued from previous page}} \\
\toprule
\textbf{Condition} & \textbf{Effect} \\
\midrule
\endhead
\bottomrule
\endfoot
Difficult Terrain & Movement costs double; Charge/Skirmish requires open ground (or Relentless talent) \\
Extreme Weather (rain, snow, heat) & First travel roll each leg is Desperate \\
Hazardous Weather (storm, blizzard, sandstorm) & Travel Timer ticks +1 at leg start; navigation DV +1 \\
Favorable Weather & First travel roll each leg gains +1 Position \\
\end{longtable}
\end{center}

\subsection{Foraging and Subsistence}
A character may forage during a leg instead of taking a role action. Roll Wits+Survival (DV by terrain, typical 3--4):
\begin{itemize}
  \item \textbf{Clean Success:} Gain 1 Supply
  \item \textbf{Partial:} Gain 1 Supply but mark 1 Fatigue (time or exposure)
  \item \textbf{Miss:} No Supply; GM gains 1 SB (scarce resources, dangerous plant, predator sign)
\end{itemize}

\section{Pacing and Arc Structure}
\label{sec:pacing-arcs}

\subsection{Session Energy Management}
Aim for three ``beats'' per session:
\begin{enumerate}
  \item \textbf{Opening hook} (5--10 minutes): Re-establish stakes, remind of timers.
  \item \textbf{Rising action} (60--90 minutes): 2--3 scenes with escalating tension.
  \item \textbf{Climax or cliffhanger} (20--30 minutes): A major roll, a timer filling, or a hard choice.
\end{enumerate}

\subsection{Arc Length by Tier}
\begin{itemize}
  \item \textbf{Tier I--II (0--90 XP)}: 3--5 session arcs. Local scope, personal stakes.
  \item \textbf{Tier III (91--150 XP)}: 6--10 session arcs. Regional influence, branching paths.
  \item \textbf{Tiers IV--V (151+ XP)}: 15+ session arcs. Continental scope, world-shaping consequences.
\end{itemize}

\subsection{Between Sessions: The GM's Sacred Trust}
\label{sec:between-sessions}
\index{GM responsibilities}

Between sessions, update timers, track complication debt, and prepare scenes. Use the \textit{Between-Session Checklist} in the GM's Appendix. Remember: your preparation is the quiet magic that transforms random encounters into meaningful episodes.

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=Between-Session Checklist]
\begin{enumerate}
  \item \textbf{Update timers:} Tick any faction, travel, or campaign timers based on last session's outcomes.
  \item \textbf{Check complication debt:} Note any unresolved player Complications; they add +1 SB to the next session's first scene.
  \item \textbf{Prepare 2--3 flexible scenes:} Not scripts, but situations with a clear stake and a timer.
  \item \textbf{Review player goals:} Look at Bonds, Patron obligations, and personal arcs. Seed at least one hook per player.
  \item \textbf{Refresh your memory:} Re-read the session summary from last time (ask a player to recap if you didn't write one).
\end{enumerate}
\end{tcolorbox}

\subsection{The Three-Timer Rule}
At any given time, maintain at most three active timers in a scene:
\begin{enumerate}
  \item \textbf{Situation Timer:} Immediate goal (persuasion, escape, research)
  \item \textbf{Counter-Clock:} Opposition pressure (guards arriving, ritual completing)
  \item \textbf{Campaign Timer:} Long-term front (faction rise, curse spreading)
\end{enumerate}
Merge or retire additional timers. Timers are visible; name them and tick them in front of players.

\section{The Finale: Calling or Forcing the Crown}
\label{sec:calling-crown}
\index{Crown}\index{Finale}

The campaign reaches its crescendo when one of two thresholds is met---the moment when accumulated influence and mounting pressure collide in a final reckoning.

\begin{tcolorbox}[colback=white!97!gray, colframe=black!80!gray, title=Finale Triggers and Conditions]
\begin{tabularx}{\linewidth}{lX}
\toprule
\textbf{Finale Type} & \textbf{Conditions and Narrative Implications} \\
\midrule
Player-Called & Mandate $\geq$ 6 and Crisis $\leq$ 3 -- the party has earned the right to choose their moment of triumph. \\
Forced Finale & Crisis $\geq$ 6 regardless of Mandate -- the world forces a confrontation that can no longer be avoided. \\
Balanced Finale & Both dials at 4--5 -- a tense equilibrium where victory and defeat hang in perfect balance. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\subsection{The Finale Procedure}
\label{subsec:finale-procedure}

When the Crown is called, run the three-beat finale:
\begin{enumerate}
  \item \textbf{Reckoning}: Defend or sanctify the record of accomplishments. Draw upon the Rival's established motives. Place the Pressure rail that will drive the scene forward.
  \item \textbf{Crossing}: Stage the kinetic rail (Escape/Hunt/Hazard) that threatens to end the scene prematurely if not managed carefully.
  \item \textbf{Coronation}: Use the Diamond Leverage to sign, seal, or swear the oath that cements the campaign's legacy.
\end{enumerate}

\subsection{Twist Collision (Finale Clause)}
\label{subsec:twist-collision}

Exactly once, when the Rival's Spade Twist contradicts their Club Belief, the table chooses:
\begin{itemize}
  \item GM gains +1 SB to complicate matters, or
  \item Players reduce two ticks total across the active rails, gaining breathing room.
\end{itemize}

\section{Legacy Conversion: Epilogue}
\label{sec:legacy-conversion}
\index{Legacy Conversion}\index{epilogue}

After the Finale, each PC draws 2 cards and answers epilogue prompts by suit. Then convert campaign elements into lasting legacy:
\begin{itemize}
  \item \textbf{Major Asset $\rightarrow$ Institution} (12 XP): A safehouse becomes a school, a spy ring becomes an intelligence service.
  \item \textbf{Seasonal Endorsement $\rightarrow$ Doctrine Rider} (4 XP): Temporary support becomes permanent policy.
  \item \textbf{Follower (Cap 3+) $\rightarrow$ Stationed NPC} (0 XP): Loyal companions become custodians of the new order.
  \item \textbf{Rival $\rightarrow$ Fixture}: Surviving adversaries become recurring elements of the setting's fabric.
\end{itemize}

\subsection{The Legacy Bond}
\label{subsec:legacy-bond}

When a PC retires or dies, their successor (a follower or a new character) inherits a \textbf{Legacy Bond}. This bond represents the predecessor's enduring influence. Once per session, the player may invoke the Legacy Bond to gain \textbf{1 Boon} when acting to uphold the predecessor's values or complete their unfinished business.

\subsection{Character Arc Management}
\label{subsec:character-arcs}

Help players develop meaningful character growth through the legacy process:
\begin{enumerate}
  \item \textbf{Establishment}: Define the character's current state and potential conflicts.
  \item \textbf{Development}: Create opportunities for growth and choice.
  \item \textbf{Crisis}: Present challenges that test the character's core beliefs.
  \item \textbf{Resolution}: Allow meaningful transformation based on choices.
\end{enumerate}

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title={Arcs Generate Obligation and Patron Interest}]
  \index{character arcs!obligation}\index{Patrons!in character arcs}
  A character's personal arc should not exist in a vacuum. Tie it directly to the game's resource systems:
  \begin{itemize}
    \item \textbf{Internal arcs} (fear, doubt, guilt) are perfect sources of \textit{Corruption} or \textit{Obligation}.
    \item \textbf{External arcs} (revenge, ambition, love) create \textit{Debt Timers} with terrestrial patrons or generate \textit{Crisis}.
    \item \textbf{Whenever a player advances their arc, mark +1 Obligation to a relevant Patron or tick a Patron's Interest timer.}
  \end{itemize}
\end{tcolorbox}

\section{Quick Start Campaigns}
\label{sec:quick-start-campaigns}

\subsection{The Sunstone Gambit (Tier I--II, 4--6 sessions)}
A medium-difficulty introduction set in Acasia.  
\textbf{Premise:} A scholar hires the party to retrieve a sunstone from a ruined tower before a rival adventuring company does.  
\textbf{Key NPCs:} Scholar Mira (quest-giver), Captain Voss (rival leader), the Tower's Guardian (ancient spirit).  
\textbf{Timers:} \emph{Rival Progress [6]}, \emph{Tower Collapse [8]}, \emph{Spirit Wakening [4]}.  
\textbf{Session breakdown:} Commission $\rightarrow$ Travel $\rightarrow$ Exploration $\rightarrow$ Confrontation $\rightarrow$ Climax $\rightarrow$ Aftermath.

\subsection{The Dying Sun (Tier III--IV, 6--8 sessions)}
A regional crisis campaign. The sun-spirit is fading; a cult of eternal night seeks to extinguish it permanently.  
\textbf{Premise:} The party must rally factions, recover a lost ritual, and confront the cult before the winter solstice.  
\textbf{Key NPCs:} High Priestess of the Sun, Cult Shadow-bound, Reluctant Ykrul Chieftain.  
\textbf{Timers:} \emph{Sun's Fading [10]} (ticks each session), \emph{Cult Influence [6]}, \emph{Alliance of the Dawn [8]}.  
\textbf{Acts:} Investigate the cult $\rightarrow$ Gather allies $\rightarrow$ Raid the cult's lair $\rightarrow$ Final confrontation.

\section{Campaign Worksheet}
\label{sec:worksheet}

Use this one-page worksheet to capture your campaign's key elements.

\begin{tcolorbox}[colback=white!97!gray, colframe=black!80!gray, title=Campaign Worksheet]
\begin{tabular}{p{4cm}p{10cm}}
\textbf{Campaign Name:} & \hrulefill \\
\textbf{Starting Tier:} & \hrulefill \quad \textbf{Estimated Sessions:} \hrulefill \\
\end{tabular}
\vspace{2mm}
\textbf{Crown Spread:} \\
\begin{tabular}{ll}
Spade (Site): & \hrulefill \\
Heart (Rival): & \hrulefill \\
Club (Pressure): & \hrulefill \\
Diamond (Leverage): & \hrulefill \\
Wild (Hidden): & \hrulefill \\
\end{tabular}
\vspace{2mm}
\textbf{Factions (name, agenda timer, key NPCs):} \\
\hrulefill \\
\hrulefill \\
\vspace{2mm}
\textbf{Fronts (name, agenda timer, supporting timers, stakes):} \\
\hrulefill \\
\hrulefill \\
\vspace{2mm}
\textbf{Mandate:} 0/6 \quad \textbf{Crisis:} 0/6 \\
\vspace{2mm}
\textbf{Player Hooks (from sheets):} \\
\hrulefill \\
\vspace{2mm}
\textbf{Notes:} \hrulefill
\end{tcolorbox}

\section{Narrative First: The World Remembers}
\label{sec:narrative-first}
\index{narrative first}

Campaign design in Fate's Edge is not about railroading players along predetermined paths---it's about \textbf{responding to player choices} with consequences that accumulate like stones in a river, gradually shaping the flow of the narrative itself. Let the world shift in response to their actions. Let factions rise and fall based on their allegiances. Let the dice sing the song of a universe that reacts.

And when the Crown is finally crowned---when the last card is played and the final timer ticks to completion---let the echo of that moment be heard across the entire Amaranthine, a testament to stories well-lived and consequences fully earned.

Remember: Your preparation between sessions is the quiet magic that transforms random encounters into meaningful episodes and mechanical challenges into memorable stories. The investment in this sacred trust pays dividends in player engagement, narrative coherence, and the creation of campaigns that will be remembered long after the final dice have been rolled.