\chapter{The Known World: A Lore Compendium}
\label{ch:lore-compendium}

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Using This Compendium]
\index{lore compendium!use}
\textbf{You only need to read the region your campaign is set in.} The generators (see the separate \textit{Travel Guide}) will provide the rest.

\medskip
\textbf{Quick reference by subgenre and typical conflict:}

\small
\begin{tabularx}{\linewidth}{lXX}
\toprule
\textbf{Region} & \textbf{Fantasy Subgenre} & \textbf{Typical Conflict} \\
\midrule
Acasia & Low fantasy / mercenary realism & Feuding warlords, mercenary flips, the Curse \\
Aelaerem & Folk horror / pastoral dark & The Neighbors, clan feuds, harvest famine \\
Aeler & Industrial / political / survival horror & Resource scarcity, jurisdiction feuds, White Flood \\
Aelinnel & Fae / mathematical / courtesy horror & Geasa, Thorn Court vs. law, the Green Gate \\
Black Banners & Military / mercenary warfare & Payday delays, condotta flips, winter campaigns \\
Ecktoria & Political intrigue / legal drama & Rival claimants, bread riots, Everflame schisms \\
Linn & Viking / maritime / oath-bound raiding & Feuds, mistland fog, Thing injunctions \\
Mistlands & Folk horror / gothic & Direwood encroachment, ward-salt shortages \\
Silkstrand & Urban / mercantile / noir & Guild rivalries, dyewater curses, condotta flips \\
Theona & Celtic / folk horror / taboo-driven & No Ninth taboo, bride-peace feasts, Green Host \\
Thepyrgos & Byzantine / academic / arcane mystery & Synod edicts, iconoclast riots, witch-hunters \\
Ubral & Highland clan / reiver / honor culture & Blood feuds, guest-right, dwarf toll roads \\
Valewood & Fairy tale / dark forest / fey remnant & Phasing ruins, fae courts, imperial echoes \\
Vhasia & Courtly romance / chivalric / fractured kingdom & Rival dukes, chevauchée, parlement war \\
Viterra & Legal drama / low fantasy / frontier justice & Border disputes, legal duels, Queen's progress \\
Ykrul & Steppe nomad / saga / honor culture & Feuds, hostage strings, white squalls \\
Zakov & Pirate / noir / crime syndicate & Syndicate wars, tide-driven routes, cursed cargoes \\
\bottomrule
\end{tabularx}
\normalsize

\medskip
When you read a region entry, look for the \textbf{GM Seed} -- a one-line hook to spark an adventure -- and the \textbf{Faction Standing} suggestion. Use the simple \(-3\) to \(+3\) scale to track attitudes, and tie advancements to \textbf{Mandate} and \textbf{Crisis} timers (see \ref{subsec:mandate} and \ref{subsec:crisis}).
\end{tcolorbox}

This chapter presents a practical overview of the major regions of Fate's Edge. For each area, you will find its dominant \textbf{fantasy subgenre}, the \textbf{primary conflicts} that drive adventure, and the \textbf{key stakeholders} (factions, leaders, and powers) that shape events. Use this as a quick reference when planning sessions or improvising.

\section{A Note on Subgenres and Travel}

Each region is designed as a distinct fantasy subgenre (see \S\ref{ch:world-interaction}). When the party travels from one region to another, the genre shifts -- a journey from the folk horror of the Mistlands to the political intrigue of Ecktoria should feel like turning the page to a different kind of story. Use the travel rules to blend these genres organically.

\section{Acasia --- \textquotedbl The Broken Province\textquotedbl}
\index{Acasia}
\textbf{Genre:} Low fantasy / mercenary realism / border feud.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Feuding petty kingdoms and warlords.
  \item Mercenary companies (the Black Banners) that shift loyalties with coin.
  \item The lingering \textquotedbl Curse\textquotedbl -- a supernatural blight that twists roads and ruins.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Margravine of the Broken March} -- The highest authority, half-saint, half-tax-roll.
  \item \textbf{The Lame King} -- A traveling pretender who makes law by pointing.
  \item \textbf{Blackwood Matriarchs} -- Hedge-witches and feud-keepers.
  \item \textbf{Free Company Captains} -- Pike and shot mercenaries between contracts.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl In Acasia, every bargain is an echo of a promise once broken, and every drink tastes faintly of debt.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A broken bridge, a cursed child, and a free company that switches sides.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Margravine secretly funds the Lame King's pretensions to keep the province fractured -- unity would threaten her tax base.
  \item A Blackwood Matriarch knows the true cause of the Curse: a forgotten oath to a powerful spirit, now demanding a child as payment.
\end{itemize}

\paragraph{Faction Standing.}
Track attitudes toward the Margravine, the Lame King, and the major free companies on a \(-3\) to \(+3\) scale. Tick \textit{Mandate} up when the party brokers a truce or earns a noble's public endorsement. Tick \textit{Crisis} when a feud spills into their path or a condotta flips.

\paragraph{Lore Note.}
\emph{The road-judges of Acasia whisper that the Curse is not a blight but a slow accounting -- every unpaid tithe, every broken oath, etched into the stone. \textquotedbl Pay what you owe,\textquotedbl\ they say, \textquotedbl or the road will eat the interest.\textquotedbl}

\section{Aelaerem (Halflings) --- \textquotedbl Hearth \& Hollow\textquotedbl}
\index{Aelaerem}
\textbf{Genre:} Folk horror / pastoral quiet with dark undercurrents.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item The \textquotedbl Neighbors\textquotedbl -- fae or otherworldly powers that demand observance.
  \item Feuds between clan matriarchs that can last generations.
  \item Harvest tithes and famine pressure.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Apple-Matron} -- Hostess of harvest feasts; power sits where she pours.
  \item \textbf{The Pale Shepherd} -- Comes when lambs are born and when people go missing.
  \item \textbf{Mummers' Captain} -- Keeps a chest of masks and stricter rules than church.
  \item \textbf{Lantern-Warden} -- Trims lamps and knows which shadow is wrong.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl Don't cross the Pale Causeway at dusk, especially not if you've got a silver tooth.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A harvest feast where the ninth cup is poured, and the Hollow answers.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Apple-Matron knows that the Pale Shepherd is not a single entity -- it is a title passed down. The current Shepherd is a desperate woman trying to buy back her son's life.
  \item The Mummers' Captain's masks are not carved; they are grown from a living barrow. Each mask contains a trapped Neighbor.
\end{itemize}

\paragraph{Faction Standing.}
Track relations with the Apple-Matron's hearth, the Lantern-Wardens, and the Pale Shepherd's circuit. Tick \textit{Mandate} when the party honors a taboo or mediates a feud. Tick \textit{Crisis} when a Neighbor is disturbed or a harvest fails.

\paragraph{Lore Note.}
\emph{A Grey Wanderer's note: \textquotedbl The Aelaerem pour the ninth cup only for the Hollow. They count to eight, then stop. I have seen a child count nine. The cup did not overflow -- but the child's reflection was no longer her own.\textquotedbl}

\section{Aeler (Dwarves) --- \textquotedbl Crowns \& Under-Vaults\textquotedbl}
\index{Aeler}
\textbf{Genre:} Industrial / political / survival horror (underground).

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Resource scarcity (air, water, light, food) in sealed holds.
  \item Jurisdictional feuds -- bell-code conflicts between crown and basilica.
  \item The \textquotedbl White Flood\textquotedbl -- mountain thaw that becomes a river through routes.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Vault-Queen} -- Sovereign above and below.
  \item \textbf{High King Beneath the Peaks} -- First among crowns by ancient concord.
  \item \textbf{True Masons} -- Wanderers repairing ancient stonework.
  \item \textbf{Edgewalkers} -- Prospectors and border scanners.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl When the stone starts singing a different tune, you best listen.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A ventilation shaft collapses, and the only way out is through a sealed crypt that holds a forgotten king's secret.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The High King's health is failing; his advisors are preparing a succession that would shatter the ancient concord.
  \item The True Masons are not neutral -- some factions use them to quietly seal off rival holds.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Vault-Queen, the High King's court, and the True Masons. Tick \textit{Mandate} when the party secures a new trade route or repairs a crucial under-way. Tick \textit{Crisis} when a bell-code dispute escalates or a White Flood warning is ignored.

\paragraph{Lore Note.}
\emph{Saikou Ira's case files: \textquotedbl An Aeler tunnel-wright once told me, \textquotedbl The mountain does not forget a misplaced keystone. It waits -- sometimes a year, sometimes a century -- then settles the debt with a collapse.\textquotedbl\ I have never seen a dwarf leave a stone unturned.\textquotedbl}

\section{Aelinnel (Gnomes) --- \textquotedbl Stone, Bough, and Bright Things\textquotedbl}
\index{Aelinnel}
\textbf{Genre:} Fae / mathematical / courtesy horror.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Geasa and spoken taboos that bind without warning.
  \item Thorn Court vs. human law -- jurisdiction debates.
  \item The Green Gate -- a threshold that demands exact change in truths.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Stone Prince of Aelinnel} -- Mortal crown with fae debts properly indexed.
  \item \textbf{Lady of Thorns} -- Hawthorn court sovereign; smiles like a snare.
  \item \textbf{Green Knight} -- Duelist of paths and proofs.
  \item \textbf{Oath-Carver} -- Sets promises in quartz so they can be seen.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The law here flows like water over stone -- it finds its own path, and woe to those who try to dam it with paper.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A missing step, an unsaid name, and a door that only opens when you speak the eighth word.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Stone Prince's ledger is missing a page -- a debt owed to the Thorn Court that could collapse his reign if discovered.
  \item The Green Knight is not a single being; the title passes to whoever wins the last duel. The current Knight is a mortal who refuses to die.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Stone Prince's court, the Lady of Thorns, and the Oath-Carvers. Tick \textit{Mandate} when the party resolves a dispute without breaking geas. Tick \textit{Crisis} when a promise is shattered or the Green Gate opens at the wrong hour.

\paragraph{Lore Note.}
\emph{From the Grey Wanderer: \textquotedbl An Aelinnel bell-tender once told me, \textquotedbl We do not say \textquotedbl no\textquotedbl\ to a spirit. We say, \textquotedbl That road is closed for repairs.\textquotedbl\ The spirit nods and leaves. It is not a lie -- the road is closed. To spirits.\textquotedbl}

\section{Black Banners --- \textquotedbl Condotta \& Crowns\textquotedbl}
\index{Black Banners}
\textbf{Genre:} Military / mercenary / low fantasy warfare.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Payday delays and supply shortages.
  \item Condotta units flipping sides mid-battle.
  \item Winter campaigns and frozen battlefields.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Banner-Captain} -- Griffon standard and a ledger of noble grudges.
  \item \textbf{The Black Colonel} -- Legendary commander not seen in months, yet his banner still rides.
  \item \textbf{Clan-Mother} (Ykrul) -- Wears war-braids from oath-breakers' hair.
  \item \textbf{Veteran Sergeant} -- Counts scars like coins.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl A contract is only as good as the next payday, and a grudge is only as sharp as the sword backing it.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{Two rival companies encamped on opposite sides of a frozen river; a single bag of coin could end the war or set off a massacre.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Black Colonel is dead; his banner is carried by a mute lieutenant who follows his last orders. No one dares question the legend.
  \item The Clan-Mother is secretly selling war-braids to both sides, prolonging the conflict for profit.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with individual banner captains, the condotta brokers, and the Clan-Mother. Tick \textit{Crisis} when a payday is missed or a condotta flips mid-contract.

\paragraph{Lore Note.}
\emph{A camp whisper: \textquotedbl The Black Colonel died three winters ago. The banner rides because the company owes him a war. They will not stop until the debt is paid -- in blood or gold.\textquotedbl}

\section{Ecktoria --- \textquotedbl Marble \& Fire\textquotedbl}
\index{Ecktoria}
\textbf{Genre:} Political intrigue / legal drama / fallen empire.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Rival dukes, abbesses, and parlements claiming legitimacy.
  \item Bread riots, grain shortages, and currency devaluation.
  \item The Everflame orthodoxy vs. reformist sects.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Grand Magistrate} -- Guardian of forms, breaker of them when needed.
  \item \textbf{High Priest of the Everflame} -- Smiles like law; flame judges truth and debt.
  \item \textbf{Duchess-Regent} -- Speaks through cousins, rules through silence.
  \item \textbf{Censor's Clerk} -- \textquotedbl Misfiles\textquotedbl\ careers; knows which names to forget.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The law here is written not in books but in marble and flame.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A forged will, a missing tax ledger, and a procession that turns into a riot at noon.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Grand Magistrate suspects the High Priest of the Everflame is hiding a heretic in the basilica's crypt -- a former Censor who knows too much.
  \item The Duchess-Regent is not the real ruler; she is a body double. The true duchess died years ago, and her spymaster keeps the secret to control the province.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Grand Magistrate, the Everflame priesthood, and the ducal courts. Tick \textit{Mandate} when the party wins a legal case or brokers a grain deal. Tick \textit{Crisis} when a bread riot ignites or a heretic is burned.

\paragraph{Lore Note.}
\emph{Saikou Ira notes: \textquotedbl The Everflame's inquisitors burn heretics on the Triumph Stairs. But the ashes are swept into the same crypts as the senators' bones. In Ecktoria, even the dead cannot escape politics.\textquotedbl}

\section{Linn --- \textquotedbl Skerries \& Storm-Oaths\textquotedbl}
\index{Linn}
\textbf{Genre:} Viking / maritime / oath-bound raiding.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Feuds that turn weddings into funerals.
  \item Mistlands fog and misdirection -- channels \textquotedbl move\textquotedbl\ without warning.
  \item Thing injunctions that forbid sailing until a grievance is heard.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Sea-Queen} -- Crown of hammered silver, eyes for weather and law.
  \item \textbf{High Jarl} -- First among raids, last to boast.
  \item \textbf{Volva of the Mist} -- Oath-reader whose nod stills surf.
  \item \textbf{Skald} -- Verses buy silence, start fights, end them.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The sea keeps better counsel than any scribe, and the storm remembers every oath sworn upon its winds.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A bride stolen, a longship lost in fog, and a skald's verse that names the dead.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Sea-Queen lost an eye to a storm -- or to a rival's blade. Her court knows, but no one speaks of it.
  \item The Volva of the Mist is not Linnic; she is a runaway from the Mistlands who bargained with a spirit for her position.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Sea-Queen's court, the High Jarl's household, and the skaldic circles. Tick \textit{Mandate} when the party resolves a feud at Thing. Tick \textit{Crisis} when a ship is lost in a squall or a wedding turns to blood.

\paragraph{Lore Note.}
\emph{\textquotedbl A Linn skald's verse: \textquotedbl The Thing-bell rings for the living. The drowned wait for a different bell -- one that only sounds when the tide forgets its name.\textquotedbl}

\section{Mistlands --- \textquotedbl Bells, Salt, and Breath\textquotedbl}
\index{Mistlands}
\textbf{Genre:} Folk horror / gothic / supernatural dread.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item The Direwood's encroachment -- undead and wraiths testing the bell-lines.
  \item Ward-salt shortages and bell-line failures.
  \item Aeler protectorate authority vs. local independence.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Legate of the Mists} -- Aeler governor; equal parts mercy and decree.
  \item \textbf{Bell-Warden} -- Ringbook decides which notes keep which ghosts at bay.
  \item \textbf{Salt-Monk} -- Blesses pans and keeps ward-measure honest.
  \item \textbf{Mist-Seer} -- Bargains breath for answers; candles shorten when she speaks.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The bells here do more than mark time -- they weave the very fabric of reality.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A bell-line fails at dusk, and something that should not walk steps through the gap.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Legate of the Mists knows that the Direwood is not a natural forest -- it is a prison for a forgotten god. The bell-lines are the bars of its cage.
  \item A faction of Salt-Monks hoards pure salt to sell to desperate villages, letting other wards decay.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Legate's protectorate, the Bell-Wardens, and the Salt-Monks. Tick \textit{Mandate} when the party repairs a bell-line or negotiates a ward-salt shipment. Tick \textit{Crisis} when a wraith breaks through or a village is quarantined.

\paragraph{Lore Note.}
\emph{A Grey Wanderer's warning: \textquotedbl In the Mistlands, never breathe deeply after the ninth bell. The fog remembers your breath, and one day it will ask for it back.\textquotedbl}

\section{Silkstrand --- \textquotedbl City of Bridges \& Dyewater\textquotedbl}
\index{Silkstrand}
\textbf{Genre:} Urban / mercantile / noir.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Guild rivalries and Exchange cornering.
  \item The \textquotedbl Curse\textquotedbl -- dyewater omens and awakened curses in the canals.
  \item Condottieri flipping colors mid-contract.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{The Matron} -- Rules from her palazzo stairs with velvet and iron.
  \item \textbf{Guildmistress of Dyers} -- Stained hands, permits tighter than purse strings.
  \item \textbf{Bridge Bailiff} -- Rents stalls, sells gossip by the breath.
  \item \textbf{Night-Boat Smuggler (\textquotedbl Ravel\textquotedbl)} -- Silent oar, louder favors.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl In this city, every bridge is a ledger and every current carries coin.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A dye-permit stamped with a dead notary's seal, and a canal that runs red at midnight.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Matron's velvet gloves hide iron claws: she personally ordered the death of the last Bridge Bailiff who refused to rig a debt auction.
  \item The Guildmistress of Dyers is slowly poisoning the water to drive out smaller competitors; the \textquotedbl Curse\textquotedbl\ is a convenient cover.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Matron's office, the Dyers' Guild, and the smuggling networks. Tick \textit{Mandate} when the party brokers a trade deal or exposes a forger. Tick \textit{Crisis} when the Exchange corners a market or a condotta flips.

\paragraph{Lore Note.}
\emph{Madam Serafine's whisper: \textquotedbl The dye-water never truly washes out. A forger's hands stay stained -- not with color, but with the memory of every seal they broke. You can see it in the way their fingers curl.\textquotedbl}

\section{Theona --- \textquotedbl Three Greens, No Ninth\textquotedbl}
\index{Theona}
\textbf{Genre:} Celtic / folk horror / taboo-driven.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item The \textquotedbl No Ninth\textquotedbl\ taboo -- omissions, unspoken names, missing steps.
  \item Bride-peace feasts that can end or restart feuds.
  \item The Green Host -- spectral riders that appear when customs are broken.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Matron of Wells} -- Queen-mother in truth if not name.
  \item \textbf{Three-Isles King} -- Elected by moots; whale-bone throne.
  \item \textbf{Lady Beneath} -- Unseen queen who bargains in dreams.
  \item \textbf{Taboo-Witness} -- Geas-keeper who knows which words ruin deals.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The wells here hold more than water -- they hold the dreams and prayers of generations.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A bride-peace feast where the ninth cup is poured, and the Green Host rides before dawn.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Lady Beneath is not a spirit -- she is a mortal exile who learned to walk through dreams. She uses omens to manipulate the Three-Isles King.
  \item The Taboo-Witness is dying. She is looking for a successor to inherit her geas -- and her debts.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Matron of Wells, the Three-Isles King's moot, and the Taboo-Witness. Tick \textit{Mandate} when the party keeps the Ninth silent or brokers a marriage peace. Tick \textit{Crisis} when a taboo is broken or the Green Host is sighted.

\paragraph{Lore Note.}
\emph{A Grey Wanderer's note: \textquotedbl I attended a moot on Theona. They counted every vote, every step, every cup -- but never the ninth. When I asked why, the eldest said, \textquotedbl The ninth is not for us. It is for the ones who will come after. Let them count it.\textquotedbl}

\section{Thepyrgos --- \textquotedbl City of a Thousand Stairs\textquotedbl}
\index{Thepyrgos}
\textbf{Genre:} Byzantine / academic / arcane mystery.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Synod edicts that proscribe or permit chants and rites.
  \item Iconoclast riots between competing faiths.
  \item Witch-hunters and arcane talent conscription.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{The Archon} -- Elected, entrenched, counting bells.
  \item \textbf{Lighthouse-Patriarch} -- Reclusive seer who steers beacons in storms.
  \item \textbf{Nomophylax} -- Law-scholar whose citations close mouths.
  \item \textbf{Palkar Captain} -- Tower guards; ladder-quick and letter-shy.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl Every stair climbed is a word in the great conversation between earth and sky.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A stolen grimoire, a silent bell, and a witch-hunter who refuses to explain her orders.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Lighthouse-Patriarch is not senile; he is hiding a child with forbidden talent in the beacon tower.
  \item The Archon plans to dissolve the Synod and rule alone. He has already bribed half the Palkar captains.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Archon's office, the Synod, and the Chain-Lantern order. Tick \textit{Mandate} when the party resolves a heresy trial or sponsors a student. Tick \textit{Crisis} when an iconoclast riot erupts or a witch-hunt goes too far.

\paragraph{Lore Note.}
\emph{Saikou Ira's observation: \textquotedbl Thepyrgos's witch-hunters carry censers of lead, not brass. \textquotedbl Brass is for faith,\textquotedbl\ they say. \textquotedbl Lead is for certainty.\textquotedbl\ I have never seen a lead censer cool.\textquotedbl}

\section{Ubral --- \textquotedbl The Stone Between Spears\textquotedbl}
\index{Ubral}
\textbf{Genre:} Highland clan / reiver / honor culture.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Feuds that last generations, settled by wergild in cattle or blood.
  \item Guest-right -- breaking hospitality stains name worse than breaking a sword.
  \item Dwarf toll roads and clan independence.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Lady of Tor} -- Clan-chief in cloak and mail.
  \item \textbf{Council of Cairns} -- Greybeards and granite wills; they do not hurry.
  \item \textbf{Feud-Broker} -- Knows the weight of life in cattle and coin.
  \item \textbf{Wergild Counter} -- Keeps tallies, ends grudges, starts others.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The stone here remembers every oath sworn upon it and every blood spilled in its shadow.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A stolen herd, a broken guest-token, and a cairn that adds a new name at dawn.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Lady of Tor is dying without an heir. She plans to name a successor from outside the clan -- a move that will trigger a civil war.
  \item The Council of Cairns is not a council of elders; it is a single shaman wearing many masks, speaking with different voices.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Lady of Tor, the Council of Cairns, and the Khaz-Vurim dwarf wardens. Tick \textit{Mandate} when the party pays a wergild or brokers a truce. Tick \textit{Crisis} when a feud reignites or a guest-right is violated.

\paragraph{Lore Note.}
\emph{A Grey Wanderer's memory: \textquotedbl An Ubral widow once told me, \textquotedbl My husband's feud cost three cows and a boy's finger. I paid the cows. The finger -- that is still owed.\textquotedbl\ She smiled. I did not ask more.\textquotedbl}

\section{Valewood --- \textquotedbl Empire Under Leaves\textquotedbl}
\index{Valewood}
\textbf{Genre:} Fairy tale / dark forest / fey remnant.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Phasing ruins and star-roads that shift between worlds.
  \item Fae courts (Hazel vs. Alder) competing for influence.
  \item Imperial echoes -- ancient shades still marching patrols.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Hazel Queen} -- Lethai-ar sovereign in leaf and law.
  \item \textbf{Alder King} -- Twilit Court sovereign in shadow and debt.
  \item \textbf{The Huntsman} -- Antlers of moonlight; shortcuts owe him.
  \item \textbf{Pathweaver} -- Knots safe routes into cord.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The forest here remembers empires that crumbled to dust, and its laws are older than your kingdoms.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A star-road that opens at dusk, a fae bargain that demands a voice, and an imperial shade that knows your true name.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Hazel Queen is dying from an old curse. She is looking for a mortal to inherit her crown -- and her debt to the Alder King.
  \item The Huntsman is not loyal to either court; he is a former lover of the Hazel Queen, cursed to hunt her enemies until she lifts the spell.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Hazel Queen, the Alder King, and the Imperial Echo patrols. Tick \textit{Mandate} when the party returns a relic or respects a geas. Tick \textit{Crisis} when a court's patience expires or a star-road realigns dangerously.

\paragraph{Lore Note.}
\emph{A Lethai-ar whisper: \textquotedbl The Valewood is not a place. It is a process. Every root, every leaf, every fallen emperor -- all are being digested into something new. Walk softly, or you will join the meal.\textquotedbl}

\section{Vhasia --- \textquotedbl The Fractured Sun\textquotedbl}
\index{Vhasia}
\textbf{Genre:} Courtly romance / chivalric / fractured kingdom.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Rival dukes claiming the Sun-crown.
  \item Chevauchée -- raiding parties that burn loyalty oaths as well as barns.
  \item Parlement divided -- war by paperwork.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Two Crowns} -- Rival dukes, each crowned in rival cathedrals.
  \item \textbf{Queen-Mother} -- Weaves alliances as if they were dowries.
  \item \textbf{Last Dauphin} -- Rumor with a birthmark; armies march for whispers.
  \item \textbf{Marshal-Exile} -- Sun-court's hammer, now wielded by whoever pays.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The sun may be fractured, but our authority is absolute within our domains.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A false Dauphin, a stolen crown, and a chevauchée that burns a village that swore neutrality.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Queen-Mother secretly supports both Crowns; she wants the war to continue until every rival duke is exhausted.
  \item The Last Dauphin is a commoner recruited by the Marshal-Exile. He believes his own legend.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with each of the Two Crowns, the Queen-Mother, and the Parlement clerks. Tick \textit{Mandate} when the party brokers a truce or exposes a forgery. Tick \textit{Crisis} when a chevauchée goes unanswered or a parlement divides.

\paragraph{Lore Note.}
\emph{The Grey Wanderer: \textquotedbl In Vhasia, every castle flies two banners -- one for the king who might return, and one for the lord who actually lives there. The third banner, the one for the truth, is never raised.\textquotedbl}

\section{Viterra --- \textquotedbl The Hedge-Law Realm\textquotedbl}
\index{Viterra}
\textbf{Genre:} Legal drama / low fantasy / frontier justice.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Border disputes over hay-carts and toll rights.
  \item Legal duels -- formal combat to settle property claims.
  \item The Queen's progress -- royal authority moving through semi-autonomous duchies.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Warrior Queen} -- Crown on helm, sword at side, patient sums behind gaze.
  \item \textbf{Queen's Justiciar} -- Law on road, polite as a noose.
  \item \textbf{Fen Reeve} -- Tally-rod; speaks for dike guilds.
  \item \textbf{Master Advocate Halric} -- Silver-tongued legal warrior.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl Every hedge here marks not just a boundary but a principle -- the careful balance between order and freedom.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A boundary stone moved overnight, a legal duel at dawn, and a charter that was signed in disappearing ink.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Warrior Queen is not a warrior; she is a diplomat who won the crown by out-arguing her rivals. Her sword has never been drawn.
  \item Master Advocate Halric is a spy for a foreign power, feeding legal secrets to Vhasian exiles.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Queen's court, the Queen's Justiciars, and the Fen Reeves. Tick \textit{Mandate} when the party wins a legal case or settles a border dispute. Tick \textit{Crisis} when a legal duel turns lethal or a charter is voided.

\paragraph{Lore Note.}
\emph{A Justiciar's saying: \textquotedbl The hedge does not move. The law moves. If you wake to find your boundary stone three feet east, do not curse the neighbor -- curse the clerk who slept with the surveyor's wife.\textquotedbl}

\section{Ykrul (Orcs) --- \textquotedbl Wolf Standards, Winter Camps\textquotedbl}
\index{Ykrul}
\textbf{Genre:} Steppe nomad / saga / honor culture.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Feuds and blood-price (wergild) in cattle or horses.
  \item Hostage strings -- token kin left to secure treaties.
  \item White squalls and rasputitsa (spring melt) that seal passes.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Khatun of Ring} -- Camp-law and hostage strings.
  \item \textbf{Khagan's Kin} -- Sent to bind foedus with a look.
  \item \textbf{Sky-Speaker} -- Shaman of storms; shortcuts owe them price.
  \item \textbf{Kon'reh Masters} -- Geometry arbiters and teachers.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The steppe remembers every oath sworn beneath its endless sky, and the wolf-standards carry the honor of clans that have outlasted empires.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A hostage string cut, a white squall that blinds both sides, and a Sky-Speaker who demands a blood-price for the weather.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Khagan's Kin are not blood relatives; they are children of conquered chieftains, trained to be loyal enforcers.
  \item The Kon'reh Masters secretly manipulate the geometry of oaths to favor certain clans, maintaining a hidden balance of power.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Khatun's camp, the Khagan's court, and the Sky-Speaker's circle. Tick \textit{Mandate} when the party pays a blood-price or secures a hostage exchange. Tick \textit{Crisis} when a white squall isolates a camp or a hostage string is severed.

\paragraph{Lore Note.}
\emph{A Ykrul bone-singer's whisper: \textquotedbl The wind carries more than snow. It carries the names of the forgotten. When a horse shies at nothing, it is because a name touched its ear. Speak the name, and the horse will calm -- or run.\textquotedbl}

\section{Zakov --- \textquotedbl Salt \& Serpent\textquotedbl}
\index{Zakov}
\textbf{Genre:} Pirate / noir / crime syndicate.

\paragraph{Primary Conflicts.}
\begin{itemize}
  \item Syndicate wars and the Salt Prince's levy.
  \item Tide-driven smuggling routes that shift with the water.
  \item Cursed cargoes and fae bargains at the Crimson Docks.
\end{itemize}

\paragraph{Key Stakeholders.}
\begin{itemize}
  \item \textbf{Salt Prince} -- Nominal ruler; real power in shadows.
  \item \textbf{Pirate Queen} -- Rules the Shallows with charm and cutlass.
  \item \textbf{Kraken's Tongue} -- Oracle who speaks in riddles and tides.
  \item \textbf{Syndicate Silent} -- No names, just favors and faces.
\end{itemize}

\paragraph{Quote.}
\begin{quote}
\emph{\textquotedbl The tide here keeps better ledgers than any scribe, and the reefs remember every secret whispered above their depths.\textquotedbl}
\end{quote}

\paragraph{GM Seed.}
\begin{quote}
\emph{A cursed cargo that whispers, a tide that does not rise, and a Syndicate Silent who offers a deal with no visible cost.}
\end{quote}

\paragraph{Hidden Agendas.}
\begin{itemize}
  \item The Salt Prince is a puppet; the real ruler is the Kraken's Tongue, who uses the oracle's prestige to manipulate every syndicate.
  \item The Pirate Queen is secretly a spy for a mainland power, sent to destabilize Zakov before an invasion.
\end{itemize}

\paragraph{Faction Standing.}
Track standing with the Salt Prince's court, the Pirate Queen's flotilla, and the Syndicate Silent. Tick \textit{Crisis} when a syndicate war breaks out or a cursed cargo escapes. (Mandate rarely applies in Zakov -- reputation is measured in threats, not legitimacy.)

\paragraph{Lore Note.}
\emph{Madam Serafine's advice: \textquotedbl In Zakov, a coin that stays warm is a curse. Spend it fast, or throw it into the sea. The Salt Prince's collectors can smell interest from three docks away.\textquotedbl}

\section{Other Regions (Brief)}
\index{other regions}
For completeness, these areas appear in the comprehensive guide but are less central to the core setting:

\begin{itemize}
  \item \textbf{Dungeon -- \textquotedbl Living Infrastructure\textquotedbl} -- A mythic underworld that breathes; genre: dungeon crawl as survival horror.
  \item \textbf{The Ways Between -- \textquotedbl Spiritways \& Veilways\textquotedbl} -- Liminal paths between death and dreams; genre: metaphysical / dream-logic.
  \item \textbf{The Wilds -- \textquotedbl Roads, Ruins, and Weather\textquotedbl} -- A reskin palette for any terrain -- forest, desert, tundra, coast.
  \item \textbf{Vilikari -- \textquotedbl Laurels \& Longhouses\textquotedbl} -- A federated frontier mixing Utaran law with barbarian custom.
\end{itemize}

\section{Using This Compendium at the Table}
\index{lore compendium!use}

When preparing a session or improvising, consult the region's entry for:
\begin{itemize}
  \item \textbf{Genre} -- to set the right mood.
  \item \textbf{Primary conflicts} -- to generate immediate stakes.
  \item \textbf{Key stakeholders} -- to populate scenes with recognizable NPCs.
  \item \textbf{Quote} -- to read aloud as a flavor text.
  \item \textbf{GM Seed} -- a one-sentence hook for quick inspiration.
  \item \textbf{Hidden Agendas} -- secrets that can become campaign threads.
  \item \textbf{Faction Standing} -- track attitudes with the simple \(-3\) to \(+3\) scale, and advance \textit{Mandate} or \textit{Crisis} timers as described.
\end{itemize}

Remember that travel rules (see \S\ref{ch:world-interaction}) allow you to blend these genres as the party moves from one region to another.