\chapter*{Preface: A Guide for Guides}
\label{ch:gm-preface}
\addcontentsline{toc}{chapter}{Preface}

\textit{``The road is a ledger. Every broken wheel is a debt, every lit lamp a witness. The only question is: what are you willing to owe?''} \\
--- Dusana of the Raven Road

Welcome, Game Master. You hold a book that is not a novel to be read cover-to-cover, but a toolkit to be consulted, a set of lenses to be put on and taken off as the light changes. \textit{Fate's Edge} gives you dice, clocks, and a simple resolution loop. What it does not give you is a script. That is your job, and this guide is here to make that job easier --- not harder.

\subsection*{How This Book Is Organised}

The chapters are arranged from the essential to the optional. Do not feel you must read everything before your first session.

\begin{itemize}
    \item \textbf{\autoref{ch:gm-intro}} --- The philosophy of consequences, safety tools, and Session Zero. Read this first.
    \item \textbf{\autoref{ch:gm-core}} --- The core loop, Position, Difficulty, Story Beats, and the infamous ``three-clock guideline''. This is the engine room.
    \item \textbf{\autoref{ch:gm-resources}} --- Supply, Fatigue, Boons, Followers, Assets, and Terrestrial Patrons. Skim the first half; save the rest for later.
    \item \textbf{\autoref{ch:gm-currency-free}} --- Running a world without coin counting. The table ``When to Charge What'' (\autoref{subsec:charge-what}) is pure gold.
    \item \textbf{\autoref{ch:gm-magic}} --- Adjudicating spells, backlash, and dangerous tags. Keep the \hyperref[tab:dangerous-tags]{Dangerous TAGS table} bookmarked.
    \item \textbf{Chapters 6--10} --- Example play, survival advice, campaign clocks, high-tier play, and advanced techniques.
    \item \textbf{\autoref{ch:lore-compendium}} --- The Known World. Read only the region where your campaign starts.
    \item \textbf{\autoref{ch:gm-reference}} --- Core tables, optional rules, rival generator, and the one-page GM screen (\autoref{app:gm-screen}). Print the screen. Use the rival generator (\autoref{app:rival-generator}) when you need a quick antagonist.
\end{itemize}

\subsection*{A Reading Plan for New GMs}

You do not need to master 200 pages to run a great session. Here is what you actually need:

\begin{enumerate}
    \item Read \textbf{\autoref{ch:gm-intro}} (safety, philosophy, Session Zero).
    \item Read \textbf{\autoref{ch:gm-core}} up to \autoref{sec:core-resolution-cycle} --- the resolution cycle, Position, DV, and the Story Beat spend menu.
    \item Read \textbf{\autoref{sec:gm-as-player}} --- practical ``now what?'' advice.
    \item Print the one-page GM screen (\autoref{app:gm-screen}) and the \hyperref[tab:outcome-matrix]{Outcome Matrix} (\autoref{tab:outcome-matrix}).
    \item Keep the \hyperref[tab:dv-ladder]{Difficulty Ladder} (\autoref{tab:dv-ladder}) and \hyperref[tab:harm]{Harm levels} (\autoref{tab:harm}) at hand.
    \item Run the Quickstart adventure (``The Lantern at Dusk'') before opening any other chapter.
\end{enumerate}

Everything else --- faction turns (\autoref{sec:faction-turn}), the Crown Spread (\autoref{sec:crown-spread}), epic tier (\autoref{ch:gm-epic}), domain management --- is waiting for you when you finish your first arc. Do not try to use it all at once.

\subsection*{The 10\% Rule}

From \autoref{ch:gm-currency-free} (and echoed in \autoref{ch:gm-reference}): pick \textit{one} subsystem from this book and use it tonight. Next session, pick another. The game will not collapse if you never use the Deck of Consequences (\autoref{app:deck-magic}) or the travel clocks (\autoref{sec:gm-travel}). The game \textit{will} collapse if you try to track every clock, every faction, every threshold effect, and every player-managed module at once.

\subsection*{What This Book Is Not}

It is not a prescribed curriculum. It is not a contract. It is a set of lenses --- put them on when you need focus, take them off when they blur the story. The Golden Rule (see \autoref{ch:gm-intro}) applies to the GM as much as to the players: when in doubt, make a ruling that keeps the story moving. Default to DV 3, Position Controlled, and let the dice fall.

\subsection*{Acknowledgements}

This guide stands on the shoulders of every GM who ran a session, forgot a rule, improvised a consequence, and still made the table laugh or gasp. Thank you for doing that work. Now go light a candle, shuffle the deck, and let the road unfold.

\vspace{1em}
\noindent\textit{--- The Grey Wanderer} \\
\textit{Castra Ferrum, Year of the Iron Harvest}