\chapter{Example Play and GM Survival Guide}
\index{example play}\index{GM survival guide}

This chapter provides two practical resources:
\begin{enumerate}
  \item A complete, annotated example session of \textbf{Fate's Edge} (Tier I--II).
  \item A short \textbf{GM Survival Guide} -- quick, actionable advice for common table situations.
\end{enumerate}
All other GM guidance (session zero, safety tools, scene framing, clocks, etc.) is covered in other chapters. This chapter focuses on \textbf{seeing the system in action} and \textbf{handling the unexpected} -- while keeping you, the GM, as a fellow player at the table.

\section{The GM as Player: A Quick Refresher}
\index{GM as player}

Before diving into the example, remember: you are not a referee or a narrator robot. You are a \textbf{co-creator of drama}, entitled to surprise, delight, and even a little mischief. Your tools (SB, clocks, Threat Pool) are not punishments -- they are your way of playing the game. When you spend a Story Beat to tick a clock, you are making a move. When you describe a guard's suspicious glance, you are adding tension. Enjoy it.

\section{Example Session: The Heist at Lord Silas's Manor}
\index{example session}\index{heist}

This walkthrough follows three characters as they attempt to steal a sealed charter from a rival merchant prince. It demonstrates how scenes are framed, how pressure escalates, and how success, failure, Boons, and Story Beats work together -- all while the GM plays an active, engaged role.

\subsection*{The Table}
\begin{itemize}
  \item \textbf{Valerius} --- Ecktorian ex-legionary (Body 3, Warfare 2)
  \item \textbf{Kira} --- Vhasian agent and negotiator (Wits 3, Sway 2)
  \item \textbf{Kael} --- Dwarven geomancer (Lore 3, Geomancy 2)
  \item \textbf{GM} -- Prepared to be surprised, to spend SB creatively, and to enjoy the ride.
\end{itemize}
The GM has prepared no fixed plot --- only factions, motives, and clocks.

\subsection*{Campaign Context}
The group seeks a sealed charter proving their patron's claim to a trade route. The charter is held by \textbf{Lord Silas}, a rival merchant prince.
\begin{itemize}
  \item Silas is politically insulated.
  \item His manor is guarded but discreet.
  \item Exposure is as dangerous as failure.
\end{itemize}

\subsection*{Scene 1: Information Gathering}
\textbf{GM frames:} \textquotedbl You are in Silkstrand's lower markets at dusk. Silas's manor rises above you. Guards are visible, but relaxed.\textquotedbl\ (The GM enjoys setting the atmospheric hook.)

\noindent\textbf{Stakes:} Success = identify an entry vector; Failure = guard suspicion escalates.
\noindent\textbf{Clock introduced:} \emph{Manor Alert} [4]

\paragraph{Action.} Kira scouts using \textbf{Wits + Subterfuge}. Position: Controlled.  
She rolls 2 Successes vs DV 3 → \textbf{Partial Success}.

\paragraph{Resolution.} GM rules:
\begin{itemize}
  \item Kira identifies a servant gate (progress achieved).
  \item GM advances \emph{Manor Alert} by 1 (cost).
  \item Kira gains \textbf{1 Boon}.
\end{itemize}
The fiction advances; pressure is visible. The GM is happy -- a Partial is the most interesting outcome.

\subsection*{Scene 2: Entry}
The group scales the outer wall at night. Position: Desperate.  
\textbf{Stakes:} Success = inside undetected; Failure = immediate pursuit.

\paragraph{Action.} Valerius rolls \textbf{Body + Athletics}.  
He rolls 4 Successes including a 10 → \textbf{Clean Success with flourish}.

\paragraph{Resolution.}
\begin{itemize}
  \item The group enters cleanly.
  \item GM reduces future Position penalties inside.
  \item No Story Beat is spent.
\end{itemize}
Strong rolls relieve pressure without removing future danger. The GM celebrates the player's cool moment.

\subsection*{Scene 3: Complication Inside}
Kael attunes to the stone to locate the vault. Position: Controlled.  
Kael rolls 0 Successes → \textbf{Miss}.

\paragraph{Resolution.}
\begin{itemize}
  \item Kael gains \textbf{2 Boons}.
  \item GM spends 1 SB to introduce a servant unexpectedly entering.
  \item GM advances \emph{Manor Alert} by 1.
\end{itemize}
The failure creates momentum rather than stalling play. The GM smiles -- now they get to describe the servant's wide eyes.

\subsection*{Scene 4: Social Pivot}
Kira calms the servant. Position: Risky.  
New clock: \emph{Servant Panic} [4].  
She rolls a \textbf{Success with SB}.

\paragraph{Resolution.}
\begin{itemize}
  \item The servant is convinced.
  \item GM spends SB to advance \emph{Servant Panic} by 2.
  \item The group now operates under time pressure.
\end{itemize}
The GM is now thinking: \textquotedbl Will they make it before the servant screams?\textquotedbl

\subsection*{Scene Cut}
The GM cuts immediately: \textquotedbl You have the vault location. You also hear boots on marble.\textquotedbl  
This prevents overplaying and preserves tension. The GM enjoys the players' anxious glances.

\subsection*{Scene 5: The Vault}
The vault is magically reinforced. Clock: \emph{Vault Integrity} [6].  
Kael attempts a geomantic bypass. He rolls Partial (2 Successes vs DV 4).

\paragraph{Resolution.} GM chooses proportional progress:
\begin{itemize}
  \item \emph{Vault Integrity} advances by 2.
  \item GM takes a Story Beat instead of increasing Alert.
  \item Kael gains 1 Boon.
\end{itemize}
Partials are \emph{GM-shaped outcomes}. The GM feels creative.

\subsection*{Scene 6: Extraction}
The vault opens as guards arrive. \emph{Manor Alert} at 3/4.  
Valerius creates a distraction while Kira escapes with the charter. The final roll succeeds --- but the GM fills \emph{Manor Alert} as the cost.

\paragraph{Outcome.}
\begin{itemize}
  \item Charter secured.
  \item Silas knows he was robbed.
  \item New faction clock begins: \emph{Silas's Retaliation} [6].
\end{itemize}
The GM is already planning the next session's opening scene.

\subsection*{Session Close}
\noindent\textbf{End-of-session review:}
\begin{itemize}
  \item Boons generated primarily from Partials and Misses.
  \item Story Beats spent to escalate, not negate success.
  \item New clocks seeded future conflict.
\end{itemize}
The session ends with a clear win, lasting consequences, and visible momentum. The GM is satisfied -- they were surprised, they played, and the table had fun.

\section{GM Survival Guide}
\index{survival guide}\index{GM tips}

This section offers quick, actionable advice for the most common GM challenges. It assumes you have read the core rules and other chapters -- this is the \textquotedbl cheat sheet\textquotedbl\ for the table.

\subsection*{When Players Go Off the Rails}
\begin{itemize}
  \item Do not panic. Ask: \textquotedbl What is your goal?\textquotedbl\ (not \textquotedbl How do you do that?\textquotedbl).
  \item Name the simplest obstacle in their way -- then set Position and DV.
  \item If they ignore the obvious, let the world react. A missed roll means the guard they bypassed earlier shows up.
  \item Use \textbf{leading questions} to steer: \textquotedbl Are you trying to sneak past or create a distraction?\textquotedbl\ instead of \textquotedbl What do you do?\textquotedbl
  \item \textbf{GM as player:} Enjoy their creativity. Your job is to say \textquotedbl yes, but\textquotedbl\ and \textquotedbl no, but\textquotedbl -- not to block them.
\end{itemize}

\subsection*{When Players Hoard Boons}
\index{boons!hoarding}
\begin{itemize}
  \item Remind them of the \textbf{Session 0 contract}: \textquotedbl Boons are not XP tokens -- they are dramatic fuel. Spend them to shape the story.\textquotedbl
  \item Introduce \textbf{time pressure} that forces spending. \textquotedbl The ritual completes in three exchanges. You have 2 Boons -- do you re-roll that failed defense or save them?\textquotedbl
  \item Use the environment: a collapsing bridge, a spreading fire, a closing gate. \textquotedbl You can improve your Position to Dominant for 1 Boon -- or risk a fall.\textquotedbl
  \item Offer a \textbf{Devil's Bargain}: \textquotedbl Spend 1 Boon now to automatically succeed, but the rival will get a head start.\textquotedbl
  \item \textbf{GM as player:} Don't punish hoarding -- make spending tempting and visible.
\end{itemize}

\subsection*{How to Make Patrons Feel Present Without Overwhelming}
\index{Patrons!GM guidance}
\begin{itemize}
  \item Use \textbf{omens} before direct demands. A flickering candle, a dropped cup, a sudden chill.
  \item Send \textbf{dreams} or visions during rest. \textquotedbl You dream of a bell that tolls eight times -- and then silence.\textquotedbl
  \item Have Patrons act through \textbf{small favors} that cost the players little but build relationship. \textquotedbl The Traveler clears the fog -- but you feel a faint tug on your heel, a reminder that roads remember footsteps.\textquotedbl
  \item Escalate slowly: first a whisper, then a request, then a demand, then a reckoning.
  \item \textbf{GM as player:} Patrons are your co-stars. Give them personality -- a lisp, a favourite phrase, a sense of humour.
\end{itemize}

\subsection*{Handling Rival Parties}
\index{rival parties!GM guidance}
\begin{itemize}
  \item Keep rivals active \textbf{off-screen}. Tick their Agenda clock when the PCs dally or fail.
  \item Use \textbf{rumours} and \textbf{found scenes} to show their progress. A burned camp, a dead informant, a signature left behind.
  \item Bring them back when players least expect it -- during a tense negotiation, at the climax of a heist, or as unexpected allies against a greater threat.
  \item Give rivals their own \textbf{Patrons} and \textbf{resources}. They should evolve just like PCs.
  \item \textbf{GM as player:} Rivals are your recurring NPC party. Enjoy running them -- they are your chance to play, too.
\end{itemize}

\subsection*{When a Scene Stalls}
\begin{itemize}
  \item Ask a leading question: \textquotedbl Who are you trying to impress?\textquotedbl\ or \textquotedbl What are you afraid of here?\textquotedbl
  \item Tick a clock. \textquotedbl While you argue, the fire spreads -- \emph{Collapse [4]} advances 1.\textquotedbl
  \item Cut to another player. Shift spotlight.
  \item \textbf{GM as player:} A stalled scene is your signal to introduce a new element. Have fun with it.
\end{itemize}

\subsection*{When You Don't Know a Rule}
\begin{itemize}
  \item Make a ruling that keeps the story moving. Say: \textquotedbl We'll do it this way for now; look it up later.\textquotedbl
  \item Default to DV 3, Controlled Position, and a consequence that feels right.
  \item Trust your gut. The system is robust.
  \item \textbf{GM as player:} Uncertainty is part of the game. Embrace it -- the dice will surprise you.
\end{itemize}

\subsection*{How to Say \textquotedbl No\textquotedbl\ (Without Firing the Player)}
\begin{itemize}
  \item \textquotedbl No, but...\textquotedbl -- offer a harder path, a cost, or a requirement.
  \item \textquotedbl Yes, but...\textquotedbl -- let them succeed, then add a complication.
  \item \textquotedbl Let's find out\textquotedbl -- call for a roll with high DV and Desperate Position. The dice decide.
  \item \textbf{GM as player:} Saying \textquotedbl no\textquotedbl\ is boring. Saying \textquotedbl no, but\textquotedbl\ is a gift to the story.
\end{itemize}

\subsection*{Using Clocks Effectively}
\begin{itemize}
  \item Name clocks clearly: \emph{Alarm Raised [4]}, not \textquotedbl Bad Stuff [6]\textquotedbl.
  \item Only keep 1--3 clocks active per scene. More than that is noise.
  \item Tick a clock on a Miss or a Partial -- even if the player otherwise succeeds.
  \item \textbf{GM as player:} Clocks are your visible toys. Tick them with relish.
\end{itemize}

\subsection*{Managing Player Spotlight}
\begin{itemize}
  \item Rotate by environment: scholar in the library, fighter on the wall.
  \item Use Bonds to pass focus: \textquotedbl You owe Rina -- what do you do to protect her?\textquotedbl
  \item Quick check-in: \textquotedbl While that happens, what is your character doing?\textquotedbl
  \item \textbf{GM as player:} Giving everyone a moment means you get to see different parts of the story.
\end{itemize}

\subsection*{When You Feel Burnt Out}
\begin{itemize}
  \item Call a break. Even five minutes resets the table.
  \item Ask players to recap -- they will remind you of forgotten details.
  \item End the session early if needed. No game is better than a bad game.
  \item Use a published adventure or a random generator for one week. Give yourself a rest.
  \item \textbf{GM as player:} Your fun matters. If you are not enjoying it, talk to your group.
\end{itemize}

\subsection*{Final Survival Mantra}
\begin{quote}
\emph{Set the DV low. Make the world react honestly. Let the pressure do the work. And remember -- you are playing too.} \\
--- The Fate's Edge GM Mantra
\end{quote}

\section{Session Zero Checklist (Handout)}
\index{Session Zero!checklist}
\label{sec:session-zero-checklist}

Use this checklist during your Session Zero. Write down the answers where everyone can see them.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Session Zero Checklist, fonttitle=\bfseries]
\begin{enumerate}
  \item \textbf{Lines \& Veils} -- What content is off-limits (Lines) or only implied (Veils)?
  \item \textbf{Tone} -- Cinematic action? Grim survival? Political intrigue? Dark horror? Mix of several?
  \item \textbf{Campaign length} -- One-shot, short arc (3--6 sessions), or long campaign (12+ sessions)?
  \item \textbf{Character hooks} -- Each player shares one goal and one fear for their character.
  \item \textbf{Patron interest} -- Which Patrons (if any) are already watching the party?
  \item \textbf{Starting situation} -- Where does the story begin? (Use the Crown Spread or a simple opener.)
  \item \textbf{House rules} -- Any optional subsystems (Threat Pool, Deck of Consequences, etc.) you plan to use?
  \item \textbf{Scheduling \& attendance} -- How often do you meet? What happens if a player misses a session?
  \item \textbf{Player-GM contract} -- Remind the table: \textquotedbl We are co-creators, not adversaries.\textquotedbl
  \item \textbf{Resource mindset} -- \textquotedbl Boons are not XP tokens. Spend them. Embrace failure.\textquotedbl
\end{enumerate}
\end{tcolorbox}

\section{The Crown Spread and the Deck of Consequences}
\label{sec:crown-deck}

These two tools are the heart of the GM-as-player philosophy. They turn campaign setup and moment-to-moment complications into a game you play alongside your group.

\subsection{Crown Spread – Campaign Skeleton in 5 Cards}

Shuffle a standard 52‑card deck (no jokers). Draw five cards and lay them in a semicircle – the Crown. For a sixth card (the Hidden), draw an additional card.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Crown Spread Meanings, fonttitle=\bfseries]
\begin{tabularx}{\linewidth}{@{} l X @{}}
\toprule
\textbf{Suit} & \textbf{Meaning} \\
\midrule
\spade\ \textbf{Spade} & \textbf{Site} – Where destiny is decided. A ruin, a bridge, a city district, a temple. \\
\heart\ \textbf{Heart} & \textbf{Rival} – Who stands between. A person with an ember: family, former ally, rival, spurned lover. \\
\club\ \textbf{Club} & \textbf{Pressure} – What prevents peace. A resource threat: famine, plague, debt, war tax. \\
\diamond\ \textbf{Diamond} & \textbf{Leverage} – The advantage. A lost letter, a relic, a hidden passage, a witness. \\
\textbf{Wild} (any suit, drawn sixth) & \textbf{Hidden Force} – The surprise. A monster, a traitor, a disaster, a forgotten god. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\noindent
\textbf{How to Use:} For each card, ask: “How does this connect to another card?” Draw arrows between them. The highest rank sets the main Threat Clock (2–5 → 4, 6–10 → 6, J/Q/K → 8, A → 10).

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=Example Crown Spread]
\begin{itemize}
    \item 7\spade → \textit{The Ruined Watchtower} (Site)
    \item Q\heart → \textit{A cunning noble with a spy in the mercenary camp} (Rival)
    \item 9\club → \textit{A blight from the tower’s well – anyone who stays weakens} (Pressure)
    \item A\diamond → \textit{A sealed charter in the crypt granting the valley} (Leverage)
    \item 10 (any suit) → \textit{A sleeping spirit bound to the foundation} (Hidden)
\end{itemize}
The Ace of Diamonds sets the campaign’s main Threat Clock to 10 segments.
\end{tcolorbox}

\subsection{Deck of Consequences – SB Complications with Flavour}

When a player rolls a \textbf{Success with SB} (successes meet DV and at least one 1), you may convert each SB into a card draw from a standard deck. Use these cards to generate complications instead of the generic SB menu.

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=Deck of Consequences Procedure]
\begin{enumerate}
    \item Note the number of SB generated (each 1 on the dice = 1 SB).
    \item Draw a number of cards equal to the SB spent.
    \item Read the combination of suits and the highest rank to synthesise a single narrative complication.
    \item \textbf{Do not reshuffle} until the deck runs out or the session ends – this creates a natural “fate budget”.
\end{enumerate}
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Card Meanings, fonttitle=\bfseries]
\noindent
\begin{tabularx}{\linewidth}{@{} l X @{}}
\toprule
\textbf{Suit} & \textbf{Interpretation} \\
\midrule
\heart\ \textbf{Hearts} & Emotional / Social – awkward moment, scandal, betrayal, broken trust. \\
\spade\ \textbf{Spades} & Harm / Escalation – bruise, injury, collapse, death. \\
\club\ \textbf{Clubs} & Material / Cost – resource loss, gear failure, debt, ruin. \\
\diamond\ \textbf{Diamonds} & Magical / Spiritual – omen, curse, Patron attention, reality bend. \\
\bottomrule
\end{tabularx}

\vspace{0.5em}
\begin{tabularx}{\linewidth}{@{} l X @{}}
\toprule
\textbf{Rank} & \textbf{Severity} \\
\midrule
2–5 & Minor inconvenience (e.g., noise, Fatigue 1, small delay). \\
6–10 & Moderate setback (e.g., alarm raised, gear broken, position worsened). \\
J, Q, K & Major consequence (e.g., reinforcements, key gear destroyed, front clock advanced). \\
A & Severe twist (e.g., trap springs, authority arrives, scene shifts). \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\noindent
\textbf{Synthesis Example:} You spend 2 SB and draw \textbf{J\heart} (social betrayal) and \textbf{7\club} (material loss). The GM narrates: “Your cloak catches on a rusty nail (resource drain – tear a scrap of fabric). Worse, the guard recognises your face from a wanted poster – she’s your cousin. She hesitates, then shouts ‘Traitor!’ (social betrayal + harm).”

\begin{tcolorbox}[colback=black!5!white, colframe=black!50!black, title=When to Use Which Tool]
\noindent
\textbf{Low cognitive load / off night:} Use the SB menu (1 SB = minor complication, 2 SB = moderate, 3+ SB = major). \\
\textbf{Medium / routine:} Use the Deck of Consequences – draw 1–3 cards per roll, interpret quickly. \\
\textbf{High energy / climax:} Draw multiple cards and weave them into a layered, memorable twist. \\
\textbf{Fallback:} DV 3, Controlled, and a simple “yes, but” is always safe.
\end{tcolorbox}

\section{Session Walkthrough – From the GM's Perspective}
\label{sec:session-walkthrough}

This walkthrough follows a single session of a Tier II game set in Silkstrand. The party consists of three characters: Kael (a Guardian), Lyra (a Free Caster), and Riley (a Psion). The campaign's \textbf{Crown Spread} is already in play: Mira Venn (Rival), the Rust (Pressure), the Tithing Bridge (Site), the Weeping Oracle (Leverage), and the Drowned Man (Wild).

\subsection{Before the Session (Prep)}
\begin{itemize}
  \item Review clocks: \textit{Mira's Consolidation} [6] (3/6), \textit{The Rust Spreads} [8] (2/8), \textit{Oracle's Safety} [6] (1/6), \textit{Drowned Man's Attention} [4] (0/4).
  \item Note last session's cliffhanger: \textquotedbl You've tracked a stolen dye shipment to a warehouse on the canal. The water is bubbling. A Scarlets patrol is approaching.\textquotedbl
  \item Prepare one opening line and two possible complications (spend 1–2 SB immediately if needed).
\end{itemize}

\subsection{Opening Hook (5 minutes)}
\begin{quote}
\textit{\textquotedbl The warehouse door is ajar. Inside, you see crates marked with Mira's seal -- a red spider. The canal behind you bubbles louder. The patrol will be here in moments. What do you do?\textquotedbl}
\end{quote}
\textbf{GM note:} No roll yet. Frame the pressure visibly. Players declare intent.

\subsection{Scene 1 – Infiltrating the Warehouse (15 minutes)}
\textbf{Player (Kael):} \textquotedbl I sneak to the side window and peer inside. How many guards?\textquotedbl

\textbf{GM:} \textquotedbl Roll Wits + Stealth, Controlled, DV 3 to get a clear view without being seen.\textquotedbl

\textbf{Kael rolls:} 1 success (needs 3) → Miss. Gains 2 Boons. GM gains 2 SB.

\textbf{GM (spends 1 SB immediately):} \textquotedbl You see two guards -- but as you lean, a loose board creaks. One guard looks toward the window. Tick the \textit{Patrol Alert} clock [4] by 1. You have one beat to hide or act.\textquotedbl

\textbf{GM (banks 1 SB for later):} \textquotedbl Riley, you sense a cold thought -- something in the canal is watching the warehouse. It feels patient.\textquotedbl

\textbf{GM note:} Use the Miss to introduce tension, not failure. The player gained Boons; the scene escalates.

\subsection{Scene 2 – Confrontation with the Scarlets (20 minutes)}
Lyra decides to create a distraction. \textquotedbl I cast a small [WIND] [STRIKE] to knock over a stack of crates behind the guards.\textquotedbl

\textbf{GM:} \textquotedbl Controlled, DV 3. Roll Wits + Arcana.\textquotedbl

\textbf{Lyra rolls:} 3 successes (DV 3) but rolls a 1 → Success with SB. GM gains 1 SB (now banked 2). Instead of the SB menu, the GM decides to use the \textbf{Deck of Consequences}.

\textbf{GM draws 1 card:} 5\spade.

\textbf{Interpretation:} 5\spade → Minor, place/hazard. GM narrates: “The crates crash. Both guards spin around. Kael, you're now in Dominant Position to slip inside. But the spilled cargo is a pile of \textit{salt jars} – the air fills with stinging dust. Everyone in the warehouse tests Body + Endurance (DV 3) or coughs uncontrollably (mark 1 Fatigue).”

\textbf{GM (spends the remaining 1 SB from earlier):} \textquotedbl You also hear the canal water stir again. The Drowned Man's Attention clock ticks to 1/4.\textquotedbl

\subsection{Scene 3 – The Drowned Man Enters (10 minutes)}
Riley: \textquotedbl I try to calm the Drowned Man with a telepathic image of still water. I want him to attack the Scarlets, not us.\textquotedbl

\textbf{GM:} \textquotedbl Desperate Position (he's a wild spirit), DV 4. Roll Spirit + Psionics. Mark 1 Mental Fatigue.\textquotedbl

\textbf{Riley rolls:} 2 successes (Partial) → gains 1 Boon. GM gains 1 SB. GM draws 1 card: \textbf{K\heart}.

\textbf{Interpretation:} K\heart → Major social consequence. GM: “The Drowned Man hesitates – his rage focuses on the guards. He climbs onto the dock and roars. The guards flee. However, your mental image leaves a residue. You now see the Drowned Man's last memory: a woman in a red coat pushing him into the canal. Her face is Mira's. You have 1 Mental Fatigue left. The spirit also imprints a fragment of its obsession on you – whenever you see the colour red, you must test Resolve (DV 3) or be distracted (–1 die on your next action).”

\subsection{Scene 4 – The Oracle's Chamber (25 minutes)}
The party finds the Weeping Oracle in a flooded basement. She speaks in riddles. The player must offer a memory to gain her knowledge.

\textbf{Kael:} \textquotedbl I offer the memory of the day my father disowned me.\textquotedbl

\textbf{GM:} \textquotedbl No roll. She drinks the memory and whispers: `The Rust is not a curse. It is Mira's dye – she poisons the water to make her red cloth unique. To stop it, you must burn her workshop on the Tithing Bridge. But the bridge is where she drowned her rival. He will not let you pass without a price.'\textquotedbl

\subsection{Scene 5 – The Tithing Bridge Climax (30 minutes)}
The party arrives at the bridge. The Drowned Man waits. The patrol has regrouped behind them. Two clocks are visible: \textit{Drowned Man's Attention} (now 1/4) and \textit{Patrol Arrives} [6] (3/6).

\textbf{GM:} \textquotedbl You're caught between. To burn the workshop, someone must cross the bridge and face the Drowned Man. Others must hold off the patrol. Who does what?\textquotedbl

The players split. Kael holds the bridge approach, Lyra and Riley cross.

\textbf{Riley:} \textquotedbl I use my last Mental Fatigue to telepathically show the Drowned Man that Mira is the woman in the red coat. I want him to let us pass.\textquotedbl

\textbf{GM:} \textquotedbl Desperate, DV 5 (convincing a spirit). Roll Spirit + Psionics.\textquotedbl

\textbf{Riley rolls:} 4 successes (Clean Success, no 1s). He spends 1 Boon to improve Position? No – already Clean.

\textbf{GM:} \textquotedbl The Drowned Man steps aside. `Go. Burn her work. Then I will rest.' Mark 1 XP for a pivotal success.\textquotedbl

Meanwhile, Kael fights the patrol. He uses his remaining Boons to re-roll a failed defense and succeeds with a Partial. The GM spends the banked 3 SB from earlier, drawing 3 cards: \textbf{2\club, 8\spade, Q\heart}.

\textbf{Synthesis:} Low material cost (2\club), moderate harm (8\spade), major social twist (Q\heart). GM: “You drive them back, but a crossbow bolt grazes your arm – mark Harm 1. One wounded guard screams that Mira will hear of this – tick \textit{Mira's Consolidation} to 4/6. Worse, the bolt had a scrap of paper tied to it: a letter to the patrol captain, signed by a member of your own patron's household. Someone in your own camp is a traitor.”

\subsection{Closing the Session (5 minutes)}
The party torches the workshop. The Drowned Man sinks into the canal, nodding. The Rust's clock stops advancing. The Oracle's Safety clock remains at 1/6 – she is safe for now.

\textbf{GM:} \textquotedbl As the fire lights the night sky, you see a silhouette on a distant rooftop – a woman in a red coat. Mira has seen you. She turns and vanishes. End of session.\textquotedbl

\subsection{Key Takeaways for GMs}
\begin{enumerate}
  \item \textbf{The Deck of Consequences turns every SB into a story.} The salt dust, the memory imprint, the traitor's letter – none of these were pre-written. They came from cards.
  \item \textbf{Clocks are promises.} Players saw the Patrol Alert clock tick and felt pressure without you saying “three rounds”.
  \item \textbf{Partial successes keep momentum.} Riley's Partial gave him the vision of Mira's face, not “you fail”.
  \item \textbf{End on a cliffhanger.} Mira's silhouette and the traitor's letter set up the next session.
\end{enumerate}

\section{The Adversary Foundry}
\label{sec:adversary-foundry}

Generate a memorable foe in 60 seconds using the following tables.

\subsection*{Archetype (d66)}

\begin{table}[h]
\centering
\small
\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
\hline
\multicolumn{2}{|c|}{\textbf{Humanoid}} \\
\hline
11-16 & Bandit / Thief \\
17-22 & Soldier / Guard \\
23-28 & Cultist / Zealot \\
29-34 & Mercenary \\
35-40 & Noble / Courtier \\
41-46 & Assassin / Spy \\
\hline
\multicolumn{2}{|c|}{\textbf{Beast}} \\
\hline
47-52 & Wolf / Pack \\
53-58 & Bear / Troll \\
59-64 & Giant Spider \\
65-70 & Serpent / Wyrm \\
71-76 & Dire Boar \\
77-82 & Roc / Giant Bird \\
\hline
\end{tabular}
\end{table}
\begin{table}[h]
\centering
\small
\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
\hline
\multicolumn{2}{|c|}{\textbf{Spirit}} \\
\hline
83-88 & Ghost / Wraith \\
89-94 & Poltergeist \\
95-100 & Echo \\
101-106 & Procession \\
107-112 & Anchor \\
113-118 & Wight \\
\hline
\multicolumn{2}{|c|}{\textbf{Cursed}} \\
\hline
119-124 & Ghoul \\
125-130 & Vampire Spawn \\
131-136 & Lycanthrope \\
137-142 & Slasher \\
143-148 & Plague Bearer \\
149-154 & Changeling \\
\hline
\end{tabular}
\end{table}
\begin{table}[h]
\centering
\small
\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
\hline
\multicolumn{2}{|c|}{\textbf{Fae}} \\
\hline
155-160 & Trow \\
161-166 & Hag \\
167-172 & Puck / Trickster \\
173-178 & Thorn Knight \\
179-184 & Green Courtier \\
185-190 & Silk Weaver \\
\hline
\multicolumn{2}{|c|}{\textbf{Other}} \\
\hline
191-196 & Construct \\
197-202 & Ooze / Slime \\
203-208 & Plant Horror \\
209-214 & Living Shadow \\
215-220 & Memory Eater \\
221-226 & Clockwork \\
\hline
\end{tabular}
\end{table}

\subsection*{Drive (d6)}
\noindent
\begin{tabular}{p{0.2\linewidth}p{0.7\linewidth}}
1 & Vengeance \\
2 & Hunger \\
3 & Duty \\
4 & Faith \\
5 & Greed \\
6 & Fear \\
\end{tabular}

\subsection*{Twist (d6)}
\noindent
\begin{tabular}{p{0.2\linewidth}p{0.7\linewidth}}
1 & Secret Identity \\
2 & Cursed \\
3 & Bound to Place \\
4 & Wounded \\
5 & Watched \\
6 & Desperate Bargain \\
\end{tabular}

\subsection*{Threat Level Scaling}

\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{TL} & \textbf{Body} & \textbf{Wits} & \textbf{Spirit} & \textbf{Presence} & \textbf{Harm} & \textbf{Skills} \\
\hline
0 (Flavor) & 1 & 1 & 1 & 1 & 2 & 0 \\
1 (Minion) & 2 & 1 & 1 & 1 & 3 & TL \\
2 (Standard) & 3 & 2 & 2 & 2 & 4 & TL+1 \\
3 (Elite) & 4 & 3 & 3 & 3 & 5 & TL+1 \\
4 (Boss) & 5 & 4 & 4 & 4 & 7 & TL+2 \\
5+ (Mythic) & 6 & 5 & 5 & 5 & 9 & TL+2 \\
\hline
\end{tabular}
\end{table}

\subsection*{Drive \& Twist Mechanical Hooks}

\begin{table}[h]
\centering
\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
\hline
\textbf{d6} & \textbf{Drive Hook} \\
\hline
1 & When damaged, gain +1 die on next attack against the source. \\
2 & Must feed (attack) each round or suffer –1 Position. \\
3 & Will sacrifice to protect a specific person, object, or place. \\
4 & Cannot be reasoned with while sacred object/shrine is present. \\
5 & Can be bribed with treasure (1 Boon or significant item). \\
6 & When source of fear is present, acts at Desperate Position. \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
\hline
\textbf{d6} & \textbf{Twist Hook} \\
\hline
1 & Hidden weakness (silver, salt, true name) – DV 3 to discover. \\
2 & Takes +1 Harm from a specific damage type. \\
3 & Cannot leave a specific area; gains +1 die within it. \\
4 & Starts with Harm 1; if healed, gains +1 die. \\
5 & A patron watches; on death, patron learns the party's location. \\
6 & Will surrender if offered resolution to core need. \\
\hline
\end{tabular}
\end{table}

\subsection*{Quick Build Example}

\noindent
\textbf{Roll:} 54 (Beast: Bear), 2 (Drive: Hunger), 5 (Twist: Watched) \\
\textbf{Frost-Touched Bear (TL2)}
\begin{itemize}
    \item \textit{Drive}: Must feed each round or suffer –1 Position.
    \item \textit{Twist}: A shaman watches through its eyes. On death, shaman learns party location.
    \item \textit{Stats}: Body 3, Wits 2, Spirit 2, Pres 1. Harm 4. Claw (Body+Melee, Harm 2).
\end{itemize}

\noindent
\textbf{Roll:} 147 (Cursed: Plague Bearer), 4 (Drive: Faith), 1 (Twist: Secret Identity) \\
\textbf{The Weeping Saint (TL3)}
\begin{itemize}
    \item \textit{Drive}: Cannot be reasoned with while her reliquary is present.
    \item \textit{Twist}: True weakness is her childhood name; speaking it grants +2 die to social rolls.
    \item \textit{Stats}: Body 2, Wits 3, Spirit 4, Pres 3. Harm 5. Contagion Aura (Near, Endurance test or Fatigue 1).
\end{itemize}

\section{The Faction Turn}
\label{sec:faction-turn}
\index{factions!turn}\index{clocks!faction}

Between sessions (or at the start of a session), you can run a quick \textbf{faction turn} to keep the world alive. A faction turn takes 5–10 minutes and answers: \textit{“What have the major powers done while the party was busy?”}

\subsection{When to Run a Faction Turn}
\begin{itemize}
  \item At the end of a session, after players declare their next destination.
  \item At the beginning of a session, as a \textquotedbl news from the world\textquotedbl\ recap.
  \item During a long downtime (several weeks or months).
\end{itemize}

\subsection{Procedure – 5 Minutes}
\begin{enumerate}
  \item \textbf{List active factions.} Keep no more than 3--5 that directly interact with the party.
  \item \textbf{For each faction, roll 1d6.} Use the table below to determine their progress.
  \item \textbf{Advance or retreat their Agenda Clock.} Each faction has a visible clock (6--10 segments) representing their goal.
  \item \textbf{Note one concrete change.} Write one sentence per faction (e.g., \textquotedbl The Scarlets seized the river docks.\textquotedbl).
  \item \textbf{Read the changes aloud} at the start of the next session.
\end{enumerate}

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, title=Faction Progress Roll (d6)]
\begin{center}
\begin{tabular}{cll}
\toprule
\textbf{Roll} & \textbf{Progress} & \textbf{Agenda Clock Change} \\
\midrule
1–2 & Setback & –1 segment (min 0) \\
3–4 & Stalled & No change \\
5–6 & Advance & +1 segment \\
\bottomrule
\end{tabular}
\end{center}
\end{tcolorbox}

\subsection{Optional: Faction Conflict Roll}
If two factions are directly opposed (e.g., the Scarlets and the Grey Circle), roll opposed d6. Higher roll wins; lower roll loses 1 segment. Ties mean a stalemate -- both lose 1 segment.

\subsection{Example Faction Turn}
\begin{quote}
\textbf{Factions:}
\begin{itemize}
  \item Mira's Scarlets – Agenda Clock: \textit{Control the Dye Trade} [8] (currently 3/8)
  \item Church of the Everflame – Agenda Clock: \textit{Suppress Heresy} [6] (currently 2/6)
  \item The Drowned Man (not a faction, but a wildcard) – no clock
\end{itemize}

\noindent \textbf{Rolls:}
\begin{itemize}
  \item Scarlets: 5 → Advance → now 4/8.
  \item Everflame: 2 → Setback → now 1/6.
\end{itemize}

\noindent \textbf{One-sentence changes:}
\begin{itemize}
  \item \textit{Scarlets:} \textquotedbl Mira's agents have bought the loyalty of the harbour master.\textquotedbl
  \item \textit{Everflame:} \textquotedbl An inquisitor was found murdered; the Church retreats to its cathedral.\textquotedbl
  \item \textit{Drowned Man:} (GM fiat) \textquotedbl The Drowned Man dragged two Scarlets into the canal. Rumours spread.\textquotedbl
\end{itemize}

\noindent \textbf{Next session opener:} \textquotedbl You wake to news that the harbour master has doubled tariffs -- and that a body was pulled from the water wearing a Scarlets' token.\textquotedbl
\end{quote}

\subsection{Using Faction Clocks in Play}
\begin{itemize}
  \item \textbf{When a clock fills}, the faction achieves its goal. The party must react (or suffer the consequences).
  \item Players can \textbf{affect faction clocks} during adventures: a successful sabotage mission might tick a clock backwards; a public failure might tick it forward.
  \item Keep faction clocks \textbf{visible} (on an index card or whiteboard). Players should see the pressure building.
\end{itemize}

\subsection{Lightweight Faction Standing (Optional)}
For factions that are not pursuing a single agenda but merely have an attitude toward the party, track a simple \textbf{-3 to +3 scale}:
\begin{itemize}
  \item \textbf{-3:} Enemy -- actively works against the party.
  \item \textbf{0:} Neutral -- indifferent.
  \item \textbf{+3:} Ally -- will sacrifice for the party.
\end{itemize}
Shift standing by ±1 after major interactions (saving a faction's leader, betraying them, etc.). No dice needed -- just fiction.

\subsection{Faction Turn Cheat Sheet (Printable)}

\begin{table}[h]
\centering
\large
\begin{tabular}{|p{0.9\linewidth}|}
\hline
\multicolumn{1}{|c|}{\textbf{THE FACTION TURN}} \\
\hline
\vspace{0.2cm}
\textbf{1. ROLL 1d6 PER FACTION} \\
\begin{tabular}{|c|c|c|}
\hline
1-2: Setback & 3-4: Stalled & 5-6: Advance \\
(-1 segment) & (0 segments) & (+1 segment) \\
\hline
\end{tabular} \\
\vspace{0.2cm}
\hline
\vspace{0.2cm}
\textbf{2. NOTE ONE CONCRETE CHANGE PER FACTION} \\
\textit{"They seized the docks." \quad "Their spy was caught." \quad "The treasury is empty."} \\
\vspace{0.2cm}
\hline
\vspace{0.2cm}
\textbf{3. RESOLVE OPPOSED FACTIONS (roll d6 vs d6)} \\
\begin{tabular}{|c|c|c|}
\hline
Winner: +1 & Loser: -1 & Tie: both -1 \\
\hline
\end{tabular} \\
\vspace{0.2cm}
\hline
\vspace{0.2cm}
\textbf{4. OPEN NEXT SESSION WITH THE RESULTS} \\
\textit{"The city is different today. You notice..."} \\
\vspace{0.2cm}
\hline
\end{tabular}
\end{table}

\subsection*{Blank Faction Tracker (Photocopy Me)}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|c|c|}
\hline
\textbf{Faction} & \textbf{Agenda} & \textbf{Clock} & \textbf{Current} \\
\hline
& & {[}8{]} & /8 \\
\hline
& & {[}6{]} & /6 \\
\hline
& & {[}6{]} & /6 \\
\hline
& & {[}4{]} & /4 \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Opposed Pairs} & \textbf{Roll Results (Winner/Loser)} \\
\hline
vs & / \\
\hline
vs & / \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{|l|}
\hline
\textbf{Next Session Opener Notes} \\
\hline
 \\
 \\
 \\
\hline
\end{tabular}
\end{table}

\section{Final GM Mantra}
\label{sec:final-mantra}

\begin{quote}
\emph{Set the DV low. Make the world react honestly. Let the pressure do the work. And remember – you are playing too. Use the Crown to begin, the Deck to surprise, the Faction Turn to breathe, and your gut to rule.}
\end{quote}