% Annotated game example (include in your docs as needed).
\section*{Annotated Master Game - The Five-Breath Weave}

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{p{2.6cm} p{5.2cm} p{5.0cm} p{3.2cm}}
\hline
\textbf{Clock} & \textbf{What it measures} & \textbf{How it ticks (±)} & \textbf{At threshold} \\
\hline
\textbf{Seed} &
Maturity of your engineered inevitability (nets, ladders, reserved captures). &
{+ place/protect anchors; force cheap replies.\newline
-- lines cut; anchors traded off.} &
High: convert; Low: delay and replant. \\
\hline
\textbf{Blue} &
Opponent’s live punish window—their fastest forcing line. &
{+ concede tempo; open lanes; mirror too long.\newline
-- close exits; trade the spearhead; change pace.} &
High: cover or counterpunch.\newline Low: invest. \\
\hline
\textbf{Ledger} &
Tempo economy—the rate at which moves pay. &
{+ double-duty moves; tolls; improving threats.\newline
-- pure defense; idle shuffles; bad trades.} &
Ahead: tax and press.\newline Behind: simplify or steal a minute. \\
\hline
\textbf{Exits} &
Health of your corridors and retreats. &
{+ fence lanes; keep two ways out; bridge files.\newline
-- starved files; pins; sealed corridors.} &
Low: reopen a file.\newline High: prosecute. \\
\hline
\textbf{Sanctum} &
Parity reserve—how many “breaths” before you must commit. &
{+ bank waiting moves; prepare equivalences.\newline
-- burn tempo; let them fix move parity.} &
Zero: zugzwang risk—trade or break symmetry. \\
\hline
\end{tabular}
\caption{The Five-Clock Weave (quick reference). Read in pairs: Seed vs. Blue, Ledger vs. Exits; Sanctum anchors both. Act when Seed is high, Blue low, Sanctum safe; stall when Blue is high and Sanctum thin.}
\end{table}


\subsection*{Ykrul (Control) vs.\ Vilikari (Tempo)}

\noindent\textbf{Players.} \textit{A (Ykrul)} “Silent Strangulation” vs.\ \textit{B (Vilikari)} “Hour-Hustle.”\\
\textbf{Clock legend:} \CC{…} Central Four timer/exclusion;\; \SC{D}/\SC{H} Blue specials spent (per life);\; \RoC{} Blue just Seeded (Rooted 1 turn);\; \RC{} Reforge countdown active;\; \CapC{a–b} on-board Greens (A–B).

\medskip
\noindent\textbf{Premises:} Second player gets an opening double-move; Twin Apex Seed fires only when Blue \emph{ends} its move on a Sanctum; Seed delay (no Seed on Blue’s \emph{first} departure); Rooted stops \emph{Blue only}; each Blue has at most \textbf{one Displacement} and \textbf{one Hop} per life.

\bigskip

\begin{multicols}{2}
\begin{enumerate}[leftmargin=*,itemsep=0.35em,label=\textbf{M\arabic*:}]

\item \textbf{A:} Red \On[L]{2} (quiet anchor). \CapC{0–0}

\item \textbf{B (double):} (i) Orange \On[R]{3} (probe);\; (ii) Red \On[L]{2} (support). \think{Tempo on day one.}

\item \textbf{A:} Second Red \On[R]{2} (mirror anchor). \think{Space now, exits later.}

\item \textbf{B:} Blue \On[R]{5} into the Central Four. \CC{in 1/3} \think{XS=2; exits to either Sanctum.}

\item \textbf{A:} Orange \Hm[L]{2} to clamp a Central Four exit. \think{Lower XS; start the squeeze.}

\item \textbf{B:} Blue \Hm[R]{4} to exit, then \textbf{Hop-capture} (land beyond on empty). \SC{H}\; \CC{out} \think{Exclusion starts.}

\item \textbf{A:} Blue \Hm[R]{4} to a Sanctum (first departure; \emph{no Seed}). \think{Mobilized; SSI still low for her.}

\item \textbf{B:} Blue to right Sanctum $\to$ \textbf{Seed} to left Sanctum. \RoC\; \CapC{0--1}

\item \textbf{A:} Declines capture; Orange \Hm[L]{2} to \emph{cover} the opposite Sanctum lane (pressures future Seed). \think{Don’t cash now—close futures.}

\item \textbf{B:} (Blue rooted) Red \On[L]{2} (center lattice). \think{Hold while \RoC{} burns.}

\item \textbf{A:} Screen: Red \On[L]{2} vs.\ Green lanes. \CapC{0--1}

\item \textbf{B:} Blue un-rooted; drifts \Hm[R]{3} toward center (threat both Sanctums).

\item \textbf{A:} Blue \Hm[L]{2} off Sanctum (set to re-end there next). \think{Parity first; pressure later.}

\item \textbf{B:} Blue \Hm[R]{4} toward a Sanctum (future Seed), avoiding a bind.

\item \textbf{A:} Blue \Hm[L]{2} back onto same Sanctum $\to$ \textbf{Seed} to opposite. \RoC\; \CapC{1–1}

\item \textbf{B:} Tempo jab: Orange \On[R]{3} to reopen a file.

\item \textbf{A:} \textbf{Hop-capture} with Blue to pick off B’s advanced Green (adjacent jump to empty). \SC{H}\; \CapC{2–0}

\item \textbf{B:} Red \On[L]{2} to print a fresh lane.

\item \textbf{A:} Not a Blue special—\textbf{Orange captures} B’s forward Red (save Displacement).

\item \textbf{B:} \textbf{Displacement} with Blue to remove A’s file plug (adjacent step). \SC{D}

\item \textbf{A:} Re-offers poison: Blue slides to a central “poison square,” adjacent to B’s Blue. \think{\CapC{2–0}; H used; D kept.}

\item \textbf{B:} Takes the bait: \textbf{captures A’s Blue} by displacement. \textbf{RC 1/5 starts for B.}

\item \textbf{A:} Killbox: Red \On[L]{2} to ZoC lane~A’s penultimate (interdict end-square).

\item \textbf{B (RC 1/5):} Orange \On[R]{3} (lane A sprint).

\item \textbf{A:} Green \Hm[R]{3} to fork lanes (blocks re-route).

\item \textbf{B (RC 2/5):} Second Orange \On[R]{3} (lane B sprint).

\item \textbf{A:} Red \On[L]{2} to stopper lane~B’s penultimate.

\item \textbf{B (RC 3/5):} Re-route; burns a ply inside ZoC.

\item \textbf{A:} Tighten ZoC with an Orange post; hold \CapC{2–0}. \think{No Seeds left; no runners spare.}

\item \textbf{B (RC 4/5):} Last angle—still two steps short into a ZoC end.

\item \textbf{A:} Trade a Red to remove the furthest sprinter.

\item \textbf{B (RC 5/5):} Fails to plant; countdown expires. \textbf{Game to A (Ykrul).}

\end{enumerate}
\end{multicols}

\paragraph{Why the Weave Works.}
A never “cashes” early value. Instead, A (i) denies future Seeds, (ii) engineers \CapC{} superiority, (iii) spends \SC{} only to prune runners, and (iv) offers a \emph{poison Blue-capture} that converts immediately into a lost \RC{} race. B’s tempo raids are excellent, but once \CapC{} reaches \texttt{2--0} and the penultimate banner squares are ZoC end-squares, every path is a step short.

\begin{coachnotes}{Coach’s Notes — Both Schools}
  \CNRow{Seed discipline}{\textbf{Seed only when SSI $\ge 2$} (their Blue distant \& specials spent or screened). If SSI < 2, you’re donating tempo.}
  \CNRow{“Free” Blue test}{\textbf{Never take a “free” Blue} unless the five-turn banner graph is solved; if you cannot plant in $\le 4$ under ZoC, the capture is a noose.}
  \CNRow{Blue specials}{Track cuts: \textbf{H}op/\textbf{D}isplacement. \textbf{1 cut left} = threaten; \textbf{0 cuts} = clamp \& Seed. Don’t spend both cuts before your safe Seed.}
  \CNRow{Cross math}{Plan exit \emph{before} entry. Enter only if you can leave cleanly next turn or force value. \CC{in 1/3} means two of \emph{your} turns remain; \CC{ex 1/2} bars you for two of \emph{your} turns.}
  \CNRow{Ring over Cross}{\textbf{Non-Blues stay out of the Cross.} Use Reds/Oranges to ring and funnel; Blue touches center briefly as a lever, not a home.}
  \CNRow{Rooted punish}{When \RoC{} shows, \textbf{post lids} (penultimate squares near Sanctums) and print lanes. Force them to defend while Blue is immobile for a breath.}
  \CNRow{Penultimate squares}{End the race here. If you control the squares one step before either Sanctum, \textbf{runners die on time}. Prefer a sturdy post over a flashy capture.}
  \CNRow{Reforge denial}{After a Blue falls, \RC{} ticks on the \emph{captor’s} turns. Deny with: (i) penultimate lids, (ii) reroute taxes (burn a ply inside ZoC), (iii) trading sprinters for time.}
  \CNRow{Poison capture}{If offered their Blue, ask: “Do I have a planted line in $\le 4$ under ZoC?” If not, \textbf{decline} or take only with an immediate lid net in place.}
  \CNRow{Green usage}{Green runs lanes, not fights. \textbf{Screen your Green} with Reds; avoid exposing it to Blue specials. Use Seed to \textbf{split attention}, then walk the quieter lane.}
  \CNRow{Mirror rule}{If you’re worse on cuts or timers, \textbf{mirror anchors then improve one file}. Don’t race a faster clock; make them waste a breath first.}
  \CNRow{Tempo trades}{Trade only when it \textbf{buys time}: a capture that opens their lane is bad; a trade that seals a penultimate square is good.}
\end{coachnotes}

{\footnotesize\textit{SSI quick read:} +1 if \CC{ex 1/2} or their Blue far; +1 if their cuts spent; +1 if your lids posted. \textbf{Seed at SSI $\ge2$.}}
