Here's the revised Player's Guide in LaTeX format, incorporating the intro and fixing the mechanical issues:

```latex
\documentclass[11pt,twoside,openany]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{titletoc}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage{makeidx}

≥ometry{letterpaper,margin=1in}
\makeindex

ℎypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Fate's Edge Player's Guide},
    pdfauthor={Nicholas A. Gasper}
}

\title{Fate's Edge: Player's Guide}
\author{Nicholas A. Gasper}
\date{October 12, 2025}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{þepage}
\fancyhead[LO]{≤ftmark}
\fancyhead[RE]{\rightmark}

\begin{document}

\maketitle

þispagestyle{empty}
\begin{center}
\textbf{Fate's Edge: Player's Guide}\\
\textcopyright\ Nicholas A. Gasper\\[1em]

This work is licensed under the Creative Commons Attribution 4.0 International License.\\
To view a copy of this license, visit:\\
\url{http://creativecommons.org/licenses/by/4.0/}
\end{center}

\tableofcontents

∈put{01ᵢntro}

\chapter{Core Mechanics} łabel{ch:core}

In this game, every action matters. The dice don't just tell you if you succeed−−−they shape the story by introducing tension, risk, and consequence. Fate's Edge is designed to keep the story moving forward, even when things go wrong. This chapter covers the core resolution system and how every roll changes the narrative.

\section*{Basic Dice Mechanics} ∈dex{dice mechanics}

When you attempt a significant action, you roll a pool of ten−sided dice (d10s). The size of your pool is determined by two factors:
$$\text{Dice Pool} = \text{Attribute} + \text{Skill}$$

\textbf{Attribute} (1−−5) Broad traits like strength, wit, or charm.

\textbf{Skill} (0−−5) Training or expertise in a specific area.

\subsection*{Reading the Dice}

Each die that rolls 6 or higher counts as a \textbf{Success}. Each die that rolls a 1 generates a \textbf{Story Beat (SB)}.

\begin{center}
\begin{tabular}{|c|l|}
ℎline
\textbf{Die Result} & \textbf{Effect} \\
ℎline
6−−10 & +1 Success \\
1 & +1 Story Beat (SB) \\
2−−5 & No effect \\
ℎline
\end{tabular}
\end{center}

\textbf{Example:} Lyra the rogue has Agility 3 and Stealth 2. Her dice pool is 5 dice. She rolls: 6, 4, 3, 1, 6. That gives her 2 Successes and 1 Story Beat. The GM sets the Difficulty Value at 2. Lyra succeeds at sneaking past the guards, but the GM now has 1 SB to spend−−−perhaps the guards hear something faintly and become suspicious.

\section*{The Description Ladder} ∈dex{description ladder}

Players can enhance their actions through detailed descriptions, which can reduce Story Beats generated by 1s:

\begin{description}
ℹtem[Basic Action] Roll the pool as−is. All 1s remain as Story Beats.
ℹtem[Detailed Action] A clear, descriptive flourish allows the player to re−roll one die showing 1.
ℹtem[Intricate Action] A richly described, multi−sensory action allows the player to re−roll all dice showing 1, and add one positive narrative flourish to the scene if they succeed.
\end{description}

\textbf{Rule:} Re−rolling 1s does not remove the Story Beats already generated by those dice. If any re−rolled dice show 1 again, they generate additional SB as normal.

\section*{Difficulty Value (DV)} ∈dex{Difficulty Value (DV)}

Before rolling, the Game Master sets a \textbf{Difficulty Value (DV)}−−−the target number of Successes needed.

\begin{center}
\begin{tabular}{|c|l|l|}
ℎline
\textbf{DV} & \textbf{Situation} & \textbf{Example} \\
ℎline
2 & Routine action, no pressure & Clear intent, modest stakes \\
3 & Pressured, mild opposition & Time pressure, partial information \\
4 & Difficult, active resistance & Hostile conditions, precise timing \\
5+ & Extreme, high stakes & Multiple constraints, dramatic failure potential \\
ℎline
\end{tabular}
\end{center}

\textbf{Tip for Players:} A DV of 3 is the most common challenge. Assume that if the GM asks you to roll, there is something at stake−−−whether it is your safety, your resources, or your reputation.

\section*{Outcome Matrix} ∈dex{outcome matrix}

Compare your Successes against the DV:

\begin{center}
\begin{tabular}{|l|l|}
ℎline
\textbf{Outcome} & \textbf{Effect} \\
ℎline
Clean Success & Goal achieved cleanly \\
Success & Cost & Goal achieved with complication \\
Partial & Progress but with difficult choice \\
Miss & No progress; complication occurs \\
ℎline
\end{tabular}
\end{center}

Let $S$ be successes and $C$ be SB generated.

\begin{description}
ℹtem[Clean Success] $(S ≥ DV \text{ and } C = 0)$ −−− Deliver the intent crisply.
ℹtem[Success & Cost] $(S ≥ DV \text{ and } C > 0)$ −−− Grant the intent; the GM spends or banks SB to add friction.
ℹtem[Partial] $(0 < S < DV)$ −−− Progress with a complication: achieve the goal with added cost, or fail forward to a different advantage.
ℹtem[Miss] $(S = 0)$ −−− No direct progress. The GM spends or banks SB to introduce immediate consequences.
\end{description}

\textbf{Player−Facing Example:} A fighter swings her sword to disarm a bandit. She rolls 3 Successes against DV 2−−−a Clean Success. The bandit's blade clatters away. Later, the same fighter tries to kick down a reinforced door with 4 dice against DV 4. She rolls only 2 Successes. This is a Partial. She cracks the door frame, but the noise attracts attention. The story moves forward either way.

\section*{Boons} ∈dex{Boons}

Boons are narrative currency that players can spend to influence the story in their favor. You can hold up to 5 Boons at a time.

\subsection*{Earning Boons}

You gain Boons through:
\begin{itemize}
ℹtem \textbf{Partial Success:} When you achieve a Partial outcome (successes $<$ DV but $> 0$), you gain 1 Boon
ℹtem \textbf{Missed Actions:} When you miss entirely (0 successes), you gain 2 Boons
ℹtem \textbf{Bond−Driven Actions:} When you take an Intricate action that meaningfully engages a character bond, you may gain 1 Boon (once per bond per session)
ℹtem \textbf{GM Award:} The GM may award Boons for creative solutions, spotlighting bonds, or meaningful sacrifices
\end{itemize}

\subsection*{Requirements for Action Awards}

Boons from Partial/miss outcomes are awarded only if:
\begin{enumerate}
ℹtem Procedure was followed correctly (intent declared, DV set, roll resolved)
ℹtem Stakes were clearly stated (what changes on success/failure)
ℹtem Consequence actually occurs (GM spends or banks SB, applies condition, or advances thread)
\end{enumerate}

\textbf{Important Note:} Rehearsal/null−risk probes and repeated identical attempts in the same scene do not award Boons. If it feels like an obvious fishing attempt, don't award a Boon.

\subsection*{Spending Boons}

You can spend Boons to:
\begin{itemize}
ℹtem Re−roll a single die in a pool
ℹtem Activate an on−screen Asset
ℹtem Power a Rite or magical ability
ℹtem Improve Position by 1 step
ℹtem Convert to XP (2 Boons = 1 XP, once per session during downtime, max 2 XP via conversion per session)
\end{itemize}

\subsection*{Carryover Limits}

At the end of each scene, reduce held Boons to a maximum of 2. Excess Boons are lost. This encourages you to spend them rather than hoard.

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Why This Matters, fonttitle=\bfseries]
The system rewards engagement with risk. Even when you don't fully succeed, you gain resources to help push the story forward. Failures become opportunities, and partial successes still offer chances to turn the tide.
\end{tcolorbox}

\section*{Story Beats (SB)} ∈dex{Story Beats (SB)}

Story Beats are narrative tools the Game Master uses to introduce twists and tension. They keep the story alive with complications and surprises.

\subsection*{What SB Can Do}

The GM may spend SB to:
\begin{itemize}
ℹtem Introduce new threats or complications
ℹtem Drain resources (time, gear, position)
ℹtem Reveal hidden dangers
ℹtem Cause collateral damage
\end{itemize}

\subsection*{SB Spend Examples}

\begin{center}
\begin{tabular}{|c|l|}
ℎline
\textbf{SB Cost} & \textbf{Effect} \\
ℎline
1 SB & Minor complication, noise, trace \\
2 SB & Moderate setback, alarm raised \\
3 SB & Serious trouble, reinforcements arrive \\
4+ SB & Major turn, scene shifts dramatically \\
ℎline
\end{tabular}
\end{center}

\textbf{Player Advice:} Don't fear Story Beats−−−they're not punishment. They are fuel for drama, ensuring the spotlight never dims.

\section*{Harm and Fatigue} ∈dex{Harm and Fatigue}

Physical injury and exhaustion are tracked through two systems:

\subsection*{Fatigue Track}

Each character has a Fatigue Track equal to their Body attribute. Mark Fatigue for:
\begin{itemize}
ℹtem Physical exertion
ℹtem Magical strain
ℹtem Travel stress
ℹtem Mental pressure
\end{itemize}

\subsection*{When your Fatigue Track fills:}
\begin{enumerate}
ℹtem Increase your Harm level by one step
ℹtem Clear all Fatigue marks
\end{enumerate}

This can happen multiple times in a scene.

\subsection*{Harm Levels}

\begin{center}
\begin{tabular}{|l|l|}
ℎline
\textbf{Harm Level} & \textbf{Effects} \\
ℎline
Harm 1 & −1 die on related actions \\
Harm 2 & −1 die on most actions until treated \\
Harm 3 & Incapacitated or dying \\
ℎline
\end{tabular}
\end{center}

\subsection*{Recovering Fatigue}
\begin{itemize}
ℹtem \textbf{Short Rest} −−− Remove 2 Fatigue with food/water
ℹtem \textbf{Full Night} −−− Remove all Fatigue
\end{itemize}

\subsection*{Recovering Harm}
\begin{itemize}
ℹtem \textbf{Minor treatment} −−− Downgrade Harm with time/rest
ℹtem \textbf{Proper medical care} −−− Remove Harm levels
ℹtem \textbf{Extended recovery} −−− Heal severe injuries
\end{itemize}

\textbf{Example:} Jorin the mercenary takes a sword cut (Harm 1). He suffers −1 die to physical actions until treated. After binding the wound and resting, the Harm fades.

\section*{Assistance} ∈dex{Assistance}

Characters can help each other. One helper per action may provide assistance by spending 1 Boon or 1 Stress, adding +1 die to the primary actor's roll. Maximum +3 dice from assists.

\textbf{Example:} Two thieves cooperate to pick a complex lock. The lead thief has Dexterity 3 + Tools 2 = 5 dice. The helper spends 1 Boon to add 1 die, making 6. Cooperation often turns failure into tense success.

\section*{Weapons & Armor} ∈dex{Weapons}

\subsection*{Weapons by Weight Class}

\begin{itemize}
ℹtem \textbf{Light (4 XP)} −−− fast, concealable.
ℹtem \textbf{Medium (8 XP)} −−− balanced, battlefield standard.
ℹtem \textbf{Heavy (12 XP)} −−− punishing, slow.
\end{itemize}

⫕section*{Melee Modifiers}

\begin{center}
\begin{tabular}{|l|c|c|l|}
ℎline
\textbf{Weight} & \textbf{Close} & \textbf{Near} & \textbf{Notes} \\
ℎline
Light & +2d & +1d & Quick, tight quarters \\
Medium & +1d & +2d & Set 1/scene or −−1d first attack \\
Heavy & −−1d & +3d & Set 1/scene or −−2d first attack \\
ℎline
\end{tabular}
\end{center}

⫕section*{Ranged & Tempo}

\begin{center}
\begin{tabular}{|l|l|l|l|l|}
ℎline
\textbf{Weight} & \textbf{Tempo} & \textbf{Close} & \textbf{Near} & \textbf{Far} \\
ℎline
Light (4 XP) & Fast & Controlled & +1d & −−− \\
Medium (8 XP) & Standard & Desperate & +2d & +1d \\
Heavy (12 XP) & Slow & Desperate & +1d & +3d \\
ℎline
\end{tabular}
\end{center}

\textbf{Tempo:} 
\begin{itemize}
ℹtem Fast = Move+Shoot. 
ℹtem Standard = Move or Shoot, Aim +1d/Effect. 
ℹtem Slow = Set/Brace, full reload, cannot Move+Shoot.
\end{itemize}

\subsection*{Weapon Tags (Optional, +4 XP each, max 2)}

Reach, Close, Accurate, Brutal, Hook, Concealable, Quickdraw, Two−Handed, Off−Hand.

\subsection*{Shields (Optional)}

\begin{center}
\begin{tabular}{|l|r|l|l|}
ℎline
\textbf{Shield} & \textbf{XP} & \textbf{Benefit} & \textbf{Tradeoff} \\
ℎline
Buckler & 4 & +1d Defend vs melee or +1 DV & Off−hand \\
Heater & 8 & +1d Defend; 1 Harm→Fatigue & −−1d Ranged \\
Pavise & 12 & Plant: heavy cover cone & Bulky, immobile \\
ℎline
\end{tabular}
\end{center}

\subsection*{Armor}

\begin{center}
\begin{tabular}{|l|r|l|l|}
ℎline
\textbf{Armor} & \textbf{XP} & \textbf{Conversion} & \textbf{Penalty} \\
ℎline
Light & 4 & 1 Harm→1 Fatigue & −−− \\
Medium & 8 & 2 Harm→1 Fatigue & −−1d physical \\
Heavy & 12 & 3 Harm→2 Fatigue & −−2d physical, no sprint \\
ℎline
\end{tabular}
\end{center}

\textbf{Notes:} Conversion applies per Harm instance before Fatigue is marked. You may still Resist first.

\subsection*{Condition & Upkeep}

\begin{description}
ℹtem[Neglected] Weapons −−1d; Armor: conversion worsens by 1 step.
ℹtem[Compromised] Weapons −−1d first attack/round; Armor: no conversion.
\end{description}

\textbf{Fix:} Short Rest/tools remove Neglected. A scene/Smith removes Compromised.

\section*{Ranged Options} ∈dex{Ranged Options}

\begin{itemize}
ℹtem \textbf{Aim:} +1d or +1 Effect.
ℹtem \textbf{Volley:} Extra ammo +1d (max +2).
ℹtem \textbf{Suppress:} Zone fire, foes −−1d/Limited Effect.
ℹtem \textbf{Overwatch:} Ready a Controlled shot on trigger.
\end{itemize}

\section*{Assets and Allies} ∈dex{Assets}

Your character's resources, contacts, or gear−−−called Assets−−−can tilt the odds in your favor.

\begin{description}
ℹtem[On−Screen Assets] −−− Companions, hirelings, or allies who stand beside you in danger
ℹtem[Off−Screen Assets] −−− Taverns, estates, titles, or networks of informants
ℹtem[Activation] −−− Spend 1 Boon to activate an on−screen Asset
\end{description}

\textbf{Narrative Use:} Assets are more than bonuses−−−they're hooks for roleplay. A friendly tavernkeeper, a noble's signet, or a trusty horse might tip the balance at the perfect moment.

\section*{Game Structure} ∈dex{Game Structure}

\subsection*{Time Scales}

\begin{description}
ℹtem[Moment] A heartbeat, a single action
ℹtem[Some Time] A few minutes, a short activity
ℹtem[Significant Time] Hours, extended effort
ℹtem[Days] Large−scale endeavors
\end{description}

\subsection*{Game Units}

\begin{description}
ℹtem[Scene] Basic narrative unit, covers specific conflict
ℹtem[Player Turn] Individual action within a scene
ℹtem[Round] Simultaneous actions in combat
ℹtem[Session] One game session (3−−6 hours)
ℹtem[Campaign] Entire story arc
\end{description}

\textbf{Player Perspective:} Think in scenes, not minutes. Every scene is a chance to shine. Every session builds toward the long arc of your campaign.

\section*{Action Resolution Steps} ∈dex{Action Resolution Steps}

\begin{enumerate}
ℹtem Describe your intent and method
ℹtem Build dice pool: Attribute + Skill (+ gear, assists)
ℹtem Roll d10s, count Successes and Story Beats
ℹtem Compare Successes to DV
ℹtem Apply outcome from matrix
ℹtem Game Master spends SB if applicable
ℹtem Earn Boons for failure.
\end{enumerate}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Quick Reference, fonttitle=\bfseries]
\textbf{Dice Pool:} Attribute + Skill d10s\\
\textbf{Success:} 6+ on each die\\
\textbf{Setback:} 1 on any die gives SB to GM\\
\textbf{DV:} 2 (easy) to 5+ (extreme)\\
\textbf{Harm:} 3−level system with penalties\\
\textbf{Boons:} 2 on miss, 1 on partial
\end{tcolorbox}

\section*{Narrative Suggestions} ∈dex{Narrative Suggestions}

\textbf{Collaborative Scene Framing:} Players may suggest scene elements (weather, NPC reactions, environmental details) that fit the established fiction, with GM approval.

\textbf{Intent−Driven Resolution:} For non−combat actions where success is reasonably assured, the GM may ask players to describe \emph{how} they accomplish their goal rather than rolling dice.

\textbf{Flashback Declarations:} Players can declare a flashback scene to establish that something happened in the past (acquiring an item, making a connection, learning information) by spending 1 Boon and describing the scene.

\textbf{Descriptive Assistance:} Players can assist each other by providing vivid, helpful descriptions of the action, granting a +1 die bonus to the primary actor's roll.

\textbf{Proactive Storytelling:} Players can suggest minor favorable details about their character's circumstances by:
\begin{itemize}
ℹtem Introducing a minor NPC who provides useful information or assistance
ℹtem Establishing that they have a useful item on hand (within reason)
ℹtem Creating a favorable environmental detail
\end{itemize}

These suggestions are subject to GM approval and should enhance rather than overshadow the main narrative.

\chapter{Character Advancement} łabel{ch:advancement}

In this game, growth isn't just about numbers−−−it's about defining who your character becomes. Advancement through Experience Points (XP) lets you shape your capabilities, influence, and legacy in the world. Every choice you make with XP is a statement about your character's priorities and the mark they leave behind.

\section*{Earning Experience Points} ∈dex{Experience Points}

XP represents learning through action. You earn it by engaging meaningfully with the world and its challenges, whether that's by triumph, failure, or bold experimentation.

\subsection*{Session Breakdown}

At the end of each session, the Game Master awards XP based on:
\begin{itemize}
ℹtem \textbf{Base Participation:} +2 XP for attending and contributing
ℹtem \textbf{Major Objectives:} +2−−4 XP for completing significant story goals
ℹtem \textbf{Discoveries:} +1−−2 XP for uncovering important lore, locations, or secrets
ℹtem \textbf{Difficult Choices:} +1−−2 XP for making hard moral or strategic decisions
ℹtem \textbf{Story Engagement:} +1−−3 XP for embracing complications and narrative twists
ℹtem \textbf{Personal Goals:} +1−−2 XP for pursuing your character's individual storylines
\end{itemize}

\textbf{Example:} At the end of a session, the party rogue earns +2 XP for participation, +2 XP for helping the group retrieve an artifact, and +1 XP for pushing a personal rivalry subplot−−−5 XP total.

\subsection*{Game Pace Options}

The GM can adjust advancement speed to match the campaign tone:

\begin{center}
\begin{tabular}{|l|l|l|}
ℎline
\textbf{Mode} & \textbf{XP/Session} & \textbf{Tone} \\
ℎline
Gritty & 4−−6 XP & Hard choices, slow growth \\
Standard & 6−−10 XP & Balanced progression \\
Epic & 10−−14 XP & Heroic, rapid development \\
ℎline
\end{tabular}
\end{center}

\textbf{Player Tip:} If you want a sweeping, mythic tale, suggest an Epic pace. For a long, hard road where each gain feels hard−earned, lean into Gritty.

\subsection*{Arc Completion Bonus}

When you finish a major story arc (typically 3−−6 sessions), everyone receives +8−−12 XP. One player may earn an additional +2 XP for a particularly memorable contribution. This celebrates the story's milestones, not just individual rolls.

\section*{Spending Experience Points} ∈dex{Experience Points!spending}

XP is your currency for growth. You can invest it in three broad areas, each representing a different approach to becoming more capable.

\subsection*{1. Personal Improvement}

Invest in your core capabilities−−−what you can do yourself.

\textbf{Attributes} Cost = New Rating × 3 XP
\begin{itemize}
ℹtem Raising Body from 2 to 3 costs 3 × 3 = 9 XP
ℹtem Raising Spirit from 4 to 5 costs 5 × 3 = 15 XP
ℹtem Requires downtime equal to new rating in days
\end{itemize}

\textbf{Skills} Cost = New Level × 2 XP
\begin{itemize}
ℹtem Improving Lore from 1 to 2 costs 2 × 2 = 4 XP
ℹtem Advancing Melee from 3 to 4 costs 4 × 2 = 8 XP
ℹtem Requires downtime equal to new level in days
\end{itemize}

\textbf{Example:} Kara wants to improve her Swordsmanship from 2 to 3. She saves 6 XP and spends three in−game days training with her mentor. This creates roleplay hooks and a sense of lived growth.

\subsection*{2. Resources and Influence}

Build your worldly presence−−−what you can command.

\begin{description}
ℹtem[Minor Resource (4 XP, 1 week)] −−− Small shop, minor contact network, basic workshop. Provides small but reliable benefits. Example: A trusted informant who gathers rumors
ℹtem[Standard Resource (8 XP, 2 weeks)] −−− Decent−sized business, skilled followers, specialized equipment. Significant benefits with some upkeep. Example: A smuggling operation with two boats
ℹtem[Major Resource (12 XP, 1 month)] −−− Large enterprise, elite team, rare capabilities. Powerful advantages with substantial upkeep. Example: A trading company with international contacts
\end{description}

\textbf{Player Tip:} Resources expand the story into new directions. A spy network creates intrigue; a workshop sparks invention; a guild hall cements influence.

\subsection*{3. Special Abilities}

Develop unique capabilities that set you apart.

\begin{description}
ℹtem[General Abilities (Cost varies)] −−− Universal benefits like improved recovery, bonus dice in specific situations, or unique combat techniques. Typically cost 4−−8 XP. Example: "Quick Recovery" − heal 1 additional Harm when resting
ℹtem[Cultural Abilities (Cost varies)] −−− Heritage−based skills tied to your character's background. Often require specific fictional positioning. Example: "Stone Sense" (dwarven) − intuitive understanding of stonework
ℹtem[Advanced Abilities (12+ XP)] −−− Powerful capstone features available at higher tiers. Often have significant narrative weight and requirements. Example: "Master Diplomat" − can reroll failed social checks once per session
\end{description}

\textbf{Example:} A veteran bard invests in "Silver Tongue" (6 XP), allowing them to sway hostile crowds once per session. This becomes their defining trick in tense negotiations.

\section*{Character Development Paths} ∈dex{Character Development Paths}

Your spending choices define your character's growth direction. Consider these archetypal paths:

\subsection*{The Specialist}

70−−90% personal improvement, 0−−10% resources, 0−−20% abilities
\begin{itemize}
ℹtem \textbf{Strengths:} Exceptional individual capability, reliable in spotlight moments
ℹtem \textbf{Weaknesses:} Limited influence, vulnerable to being isolated
ℹtem \textbf{Best for:} Solo operatives, elite warriors, master artisans
ℹtem \textbf{Example:} A duelist who invests heavily in combat skills and physical attributes
\end{itemize}

\subsection*{The Leader}

50−−65% personal, 15−−░ resources, 15−−░ abilities
\begin{itemize}
ℹtem \textbf{Strengths:} Well−rounded, can handle diverse challenges, good support
ℹtem \textbf{Weaknesses:} Jack−of−all−trades, not exceptional in any area
ℹtem \textbf{Best for:} Party faces, field commanders, investigators
ℹtem \textbf{Example:} A merchant−prince with decent combat skills, good social abilities, and a network of contacts
\end{itemize}

\subsection*{The Mastermind}

25−−40% personal, 35−−55% resources, 20−−40% abilities
\begin{itemize}
ℹtem \textbf{Strengths:} Extensive influence, can solve problems indirectly, strategic power
ℹtem \textbf{Weaknesses:} Personally vulnerable, complex upkeep, domino−effect risks
ℹtem \textbf{Best for:} Spymasters, crime lords, wealthy patrons
ℹtem \textbf{Example:} An information broker with modest personal skills but an extensive spy network
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Player Note, fonttitle=\bfseries]
These are not rigid templates. Mix and match to discover unique growth arcs.
\end{tcolorbox}

\section*{Training and Development Time} ∈dex{Training Time}

Most improvements require downtime to reflect the effort of learning and integration.

\subsection*{Standard Time Requirements}

\begin{itemize}
ℹtem Attribute increase: New rating in days
ℹtem Skill improvement: New level in days
ℹtem Resource acquisition: 1 week to 1 month depending on scope
ℹtem Ability learning: Typically 3−−10 days
\end{itemize}

\subsection*{Accelerated Development}

You can attempt to learn things more quickly, but this carries risks:
\begin{itemize}
ℹtem The GM creates a Risk Timer with 4 segments
ℹtem If the timer fills during rushed training, the new capability has flaws:
\begin{itemize}
ℹtem Attribute/Skill: −1 die penalty until you spend proper downtime
ℹtem Resource: Loyalty problems or functional limitations
ℹtem Ability: Unreliable or with unintended side effects
\end{itemize}
\end{itemize}

\textbf{Example:} The wizard crams advanced spellwork into a frantic three days. She gains the ability, but her Risk Timer fills−−−her spells now sputter unpredictably until she retrains.

\section*{Character Progression Tiers}
∈dex{Character Progression Tiers}

As you accumulate XP and capabilities, you advance through tiers that represent your growing reputation and influence.

\subsection{Tier I: Novice (0−−40 XP)}
\begin{itemize}
ℹtem Learning the ropes, establishing yourself
ℹtem Local reputation, modest capabilities
ℹtem Typical assets: Basic equipment, a few contacts
ℹtem \textbf{Position Manifestation:} The difference between life and death; a shift from Desperate to Controlled means "I live to fight another day"
\end{itemize}

\subsection{Tier II: Seasoned (41−−90 XP)}
\begin{itemize}
ℹtem Proven capability, reliable in your specialty
ℹtem Regional reputation, respected in your field
ℹtem Typical assets: Skilled followers, specialized equipment
ℹtem \textbf{Position Manifestation:} The difference between victory and defeat; a shift from Controlled to Dominant means "I've earned the right to choose how this ends"
\end{itemize}

\subsection{Tier III: Veteran (91−−150 XP)}
\begin{itemize}
ℹtem Master of your craft, significant influence
ℹtem National reputation, can handle major challenges
ℹtem Typical assets: Multiple operations, elite teams
ℹtem \textbf{Position Manifestation:} The difference between influence and control; a shift from Controlled to Dominant means "The city hangs on my decision"
\end{itemize}

\subsection{Tier IV: Paragon (151−−220 XP)}
\begin{itemize}
ℹtem Exceptional capability, strategic influence
ℹtem Continental influence, shapes regional outcomes
ℹtem Typical assets: Organizations, unique capabilities
ℹtem \textbf{Position Manifestation:} The difference between shaping events and defining eras; a shift from Controlled to Dominant means "I am the axis upon which history turns"
\end{itemize}

\subsection{Tier V: Mythic (221+ XP)}
\begin{itemize}
ℹtem Legendary status, world−changing influence
ℹtem Historical reputation, defines eras
ℹtem Typical assets: Nations, legendary artifacts
ℹtem \textbf{Position Manifestation:} The difference between ending and remaking the world; a shift from Desperate to Controlled means "The universe waits on my choice"
\end{itemize}

\subsection{Position Scales with Narrative Weight}
Position is a narrative anchor that remains consistent across tiers without mechanical scaling. The mechanical effect (re−rolling one die) never changes−−−but what it \textit{means} evolves with your tier:

\begin{itemize}
ℹtem \textbf{Tier I:} Position defines personal survival
ℹtem \textbf{Tier III:} Position shapes regional outcomes
ℹtem \textbf{Tier V:} Position reshapes reality
\end{itemize}

The power of Position is that it remains the same mechanical tool while its narrative weight grows with your influence. The system does not need to scale−−−your story does.

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!75!black, boxrule=0.5pt, arc=2mm]
\textbf{GM Tip:} Never add mechanical bonuses to Position. Instead, deepen the narrative weight:
\begin{itemize}
ℹtem \textbf{Tier I:} "Your Position shows whether you live or die"
ℹtem \textbf{Tier V:} "Your Position shows how the world bends to your will"
\end{itemize}
The mechanics stay consistent so the narrative can grow as large as your imagination allows.
\end{tcolorbox}

\section*{Managing Allies and Followers} ∈dex{Followers}

Characters who work with you require maintenance and carry risks.

\subsection*{Acquisition Costs}

\begin{itemize}
ℹtem Skilled helper: Capability rating squared in XP
ℹtem Example: A capability 3 scout costs 9 XP
\end{itemize}

\subsection*{Upkeep Requirements}

\begin{itemize}
ℹtem Each downtime period, spend XP equal to their capability rating
ℹtem Alternative: Dedicate a scene to maintaining the relationship
\end{itemize}

\section*{Strategic Advancement Considerations} ∈dex{Advancement Strategy}

\subsection*{Early Game (Tiers I−−II)}

Focus on survival and establishing your niche:
\begin{itemize}
ℹtem Invest in core competencies first
ℹtem Build a small but reliable support network
\end{itemize}

\subsection*{Mid Game (Tier III)}

Expand your influence and specialize:
\begin{itemize}
ℹtem Develop your signature capabilities
ℹtem Build substantial resources
\end{itemize}

\subsection*{Late Game (Tiers IV−−V)}

Shape the world around you:
\begin{itemize}
ℹtem Pursue advanced abilities
ℹtem Build organizations or movements
ℹtem Leave a legacy
\end{itemize}

\section*{Advancement Philosophy} ∈dex{Advancement Philosophy}

Remember that advancement serves the story. The best choices:
\begin{itemize}
ℹtem Reflect your character's experiences and growth
ℹtem Create interesting new capabilities and complications
ℹtem Enhance the group's collective abilities
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Final Thought, fonttitle=\bfseries]
Every XP spent changes not just your character sheet, but your character's story. Choose investments that make your hero more interesting to play and watch evolve.
\end{tcolorbox}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=XP Planning Guide, fonttitle=\bfseries]
\textbf{Early Tier Priorities:}
\begin{itemize}
ℹtem Core attribute to 3 (9 XP)
ℹtem Key skills to 2−−3 (4−−8 XP each)
ℹtem 1−−2 minor resources (8 XP total)
\end{itemize}

\textbf{Mid Tier Expansion:}
\begin{itemize}
ℹtem Attributes to 4 (12 XP)
ℹtem Specialization skills to 4 (8 XP)
ℹtem Standard resources (8 XP each)
ℹtem Cultural abilities (6−−10 XP)
\end{itemize}

\textbf{Late Tier Mastery:}
\begin{itemize}
ℹtem Capstone abilities (12+ XP)
ℹtem Major resources (12 XP)
ℹtem Legacy projects
\end{itemize}
\end{tcolorbox}

\section*{Narrative−Heavy Advancement Options} ∈dex{Narrative−Heavy Gameplay!Advancement}

For groups that prefer strong narrative focus in advancement, consider these optional approaches:

\textbf{Story−Driven Milestones:} Instead of tracking XP numerically, the GM can award advancement when characters reach significant story milestones. "You've trained with the master for months−−−you've improved your skill."

\textbf{Experience Through Reflection:} Players can spend downtime scenes reflecting on past experiences to earn XP. A meaningful flashback or character moment can justify growth without tracking specific points.

\textbf{Collaborative Advancement:} The group can discuss and agree on advancement choices, ensuring everyone's growth supports the overall story direction.

\textbf{Narrative Justification Focus:} When spending XP, players should explain how their character gained this capability through in−game experiences, creating richer backstory and continuity.

\chapter{Magic and Special Abilities} łabel{ch:magic}

Magic in this game is powerful but dangerous−−−a negotiation with reality itself that always carries risks. This chapter covers the core magical systems: standard spellcasting, ritual magic, and special pact−based abilities. Throughout, look for examples and player−facing tips to keep the fiction front and center.

\section*{The Nature of Magic} ∈dex{magic!nature}

Magic is not a safe tool but a dangerous force:
\begin{itemize}
ℹtem \textbf{Powerful:} Can reshape battles, stories, or even the world
ℹtem \textbf{Controlled:} Every use generates Story Beats (SB) that manifest as backlash
ℹtem \textbf{Thematic:} Effects and consequences align with the type of magic used
ℹtem \textbf{Volatile:} Never fully predictable or controllable
ℹtem \textbf{Narrative:} Casting is always a significant story moment
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Table Vignette, fonttitle=\bfseries]
"I can hold the avalanche," says Mira, fingers trembling. "But something will answer." The party nods−−−risk accepted, stakes clear.
\end{tcolorbox}

\section*{Basic Spellcasting} ∈dex{magic!spellcasting}

All spellcasting follows the standard action resolution system but with additional considerations for magical effects.

\subsection*{The Casting Process}

\begin{enumerate}
ℹtem Declare Intent: What you want the magic to achieve
ℹtem Choose Approach: Which magical skill and method you'll use
ℹtem Set Position: Dominant, Controlled, or Desperate based on circumstances
ℹtem Roll: Attribute + Magical Skill
ℹtem Resolve: Apply outcomes with magical consequences
\end{enumerate}

\subsection*{Magical Skills}

Common magical skills include:
\begin{itemize}
ℹtem \textbf{Arcana:} General magical knowledge and theory
ℹtem \textbf{Elemental Magic:} Fire, water, earth, air manipulation
ℹtem \textbf{Spiritual Magic:} Communing with spirits, divine magic
ℹtem \textbf{Mental Magic:} Telepathy, illusion, mind affecting
ℹtem \textbf{Healing Magic:} Restoration, purification, life magic
\end{itemize}

\textbf{Player Tip:} State a clear intent and a vivid method. The more concrete the fiction, the easier it is to set fair DV and meaningful consequences.

\section*{The Casting Loop} ∈dex{magic!casting loop}

For more significant magical effects, use the structured Casting Loop requiring two actions.

\subsection*{Phase 1: Weave}

Shape the magical effect:
\begin{itemize}
ℹtem Player builds dice pool and rolls
ℹtem On success, they stabilize the spell's form
ℹtem Any 1 rolled may cause narrative backlash related to the Element
\end{itemize}

\subsection*{Phase 2: Cast}

Channel the effect into the world:
\begin{itemize}
ℹtem A second roll channels the effect
ℹtem Backlash: Any 1 rolled may cause narrative backlash related to the Element
\end{itemize}

\textbf{Designer Note:} The Casting Loop requires the Spellcraft talent (6 XP) and creates spotlight tension: describe effect now, risk Backlash on each roll.

\section*{Backlash Severity} ∈dex{magic!backlash}

\begin{center}
\begin{tabular}{|l|l|}
ℎline
\textbf{Roll Result} & \textbf{Backlash Trigger} \\
ℎline
Partial/Miss & Minor backlash (choose one) \\
Miss & Major backlash (choose two) \\
Hit with two or more 1s & Minor backlash alongside success \\
ℎline
\end{tabular}
\end{center}

\section*{Magical Arts and Traditions} ∈dex{magic!traditions}

Different cultures and traditions approach magic differently.

\subsection*{Elemental Magic}

Manipulation of natural forces:
\begin{description}
ℹtem[Fire Magic:] Heat, light, transformation, destruction
ℹtem[Water Magic:] Flow, healing, divination, adaptation
ℹtem[Earth Magic:] Stability, protection, growth, strength
ℹtem[Air Magic:] Movement, communication, freedom, change
\end{description}

\subsection*{Spiritual Magic}

Interaction with intangible forces:
\begin{description}
ℹtem[Divine Magic:] Power from gods or higher powers
ℹtem[Spirit Magic:] Communing with nature spirits or ancestors
ℹtem[Necromancy:] Interaction with death and the departed
ℹtem[Protection Magic:] Wards, blessings, purification
\end{description}

\subsection*{Mental Magic}

Affecting minds and perceptions:
\begin{description}
ℹtem[Illusion:] Creating false perceptions and images
ℹtem[Telepathy:] Mind reading and communication
ℹtem[Enchantment:] Influencing thoughts and emotions
ℹtem[Divination:] Gaining knowledge through supernatural means
\end{description}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Vignette, fonttitle=\bfseries]
The candles lean toward the oracle's breath. "Ask," she whispers, "but truth is sharp."
\end{tcolorbox}

\section*{Ritual Magic} ∈dex{magic!ritual}

Rituals take Significant Time (typically 10−30 minutes) for powerful effects.

\subsection*{Ritual Requirements}

\begin{itemize}
ℹtem Time: Significant Time (typically 10−30 minutes)
ℹtem Preparation: Specific materials, locations, or conditions
ℹtem Focus: Undisturbed concentration and coordination
\end{itemize}

\subsection*{Ritual Procedure}

\begin{enumerate}
ℹtem Preparation: Gather components, prepare space, focus intent
ℹtem Invocation: Perform the Rite as a ritual
ℹtem Completion: Effect manifests, always marks +1 Obligation
\end{enumerate}

\subsection*{Ritual Benefits and Risks}

\begin{description}
ℹtem[Benefits:] Safe casting, no Push It option
ℹtem[Risks:] Time investment, Obligation cost, environmental requirements
\end{description}

\section*{Rites and Pact Magic} ∈dex{magic!rites}

Rites are precise magical effects gained through pacts with powerful entities. There are two main paths to accessing Rites:

\subsection*{The Runekeeper (Rites Path)}

\begin{itemize}
ℹtem Requires Patron + Thiasos (Familiar) + Codex (4 XP)
ℹtem Accesses that Patron's full Rite list
ℹtem Structured, powerful, but accrues Obligation
ℹtem Can Push Rites once per scene for +1 Obligation
\end{itemize}

\subsection*{The Invoker (Symbol Path)}

\begin{itemize}
ℹtem Requires one or more Patron's Symbols (4 XP each)
ℹtem Accesses ritual invocation of Patron's Rites
ℹtem Safe but slow−−−requires Significant Time
ℹtem Can Crack the Seal for instant cast at steep Obligation cost (+2/+3)
\end{itemize}

\subsection*{Using Rites}

\begin{enumerate}
ℹtem Invocation: Invoke a Rite requires 1 Action
ℹtem Obligation: Each Rite used marks Obligation on its timer
ℹtem Effect: The Rite's specific effect manifests
\end{enumerate}

\subsection*{Rite Invocation via Symbol}

\begin{itemize}
ℹtem Time. Invoking a Rite via Symbol takes DV + 1 rounds.
ℹtem Obligation. On completion, mark +1 Obligation (in addition to any listed Rite costs, if applicable).
ℹtem No Push. Invoker Rites cannot use Push It benefits.
ℹtem Symbol Display. The Symbol must remain visible throughout the invocation.
ℹtem Materials. Symbols replace any Thaisos and Codex requirements.
\end{itemize}

\section*{Obligation Capacity} ∈dex{magic!obligation}

A character's Obligation Capacity equals Spirit + Presence. Track total Obligation segments across all Patrons (or Symbols, for Invokers).

\begin{itemize}
ℹtem Exceeding Capacity: For each segment above Capacity, mark 1 Fatigue. The character cannot Invoke Rites or perform rituals until Obligation is reduced below Capacity.
ℹtem Resolution: Reduce Obligation through Downtime service, Patron tasks, ritual cleansing, or story resolution.
\end{itemize}

\textbf{Example:} Spirit 2 + Presence 3 = Capacity 5. 6 segments → Fatigue 1. 7 segments → Fatigue 2. 10 segments → Harm 1. 11 segments → Harm 2.

\subsection*{Obligation Management}

Your debt to Patrons must be managed:
\begin{itemize}
ℹtem Service: Perform tasks fitting your Patron's nature
ℹtem Offerings: Provide sacrifices or tributes
ℹtem Propagation: Spread your Patron's influence or beliefs
ℹtem Downtime: Clear through fitting service during downtime
\end{itemize}

\subsection*{Obligation Levels}

\begin{center}
\begin{tabular}{|l|l|}
ℎline
\textbf{Segments} & \textbf{Consequences} \\
ℎline
1−−2 & Minor attention, subtle signs \\
3−−5 & Noticeable influence, regular demands \\
6−−8 & Significant control, major tasks required \\
9+ & Dominant influence, potentially dangerous \\
ℎline
\end{tabular}
\end{center}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Vignette, fonttitle=\bfseries]
At the crossroads, Ash lays iron nails and salt. The wind shifts. Somewhere, something smiles.
\end{tcolorbox}

\section*{Special Magical Abilities} ∈dex{magic!special abilities}

Some characters develop unique magical capabilities through experience or heritage.

\subsection*{Cultural Magical Traditions}

\begin{description}
ℹtem[Dwarven Stone−Sense:] Intuitive understanding of earth and stone
ℹtem[Elven Memory−Weaving:] Accessing and manipulating ancestral knowledge
ℹtem[Human Versatility:] Adaptable magical approaches from various traditions
ℹtem[Nomadic Spirit−Walking:] Journeying between physical and spiritual realms
\end{description}

\subsection*{Advanced Magical Techniques}

At higher levels, casters can:
\begin{itemize}
ℹtem Specialize: Focus on specific magical traditions
ℹtem Innovate: Create new spells or techniques
ℹtem Teach: Instruct others in magical arts
ℹtem Research: Discover lost or forbidden knowledge
\end{itemize}

\section*{Magical Backlash Examples} ∈dex{magic!backlash examples}

\subsection*{Elemental Backlash}

\begin{description}
ℹtem[Fire:] Burns, flares; vs. Water: slick, sputter, dim
ℹtem[Water:] Slippery tide, slow gear; vs. Fire: smoke, shorted gear
ℹtem[Earth:] Slips, binds, encumbrance; vs. Air: sound carries, exposure
ℹtem[Air:] Scatter, misheard words; vs. Earth: stuck, dust choke
\end{description}

\subsection*{Conceptual Backlash}

\begin{description}
ℹtem[Fate:] Options close, only−one−way; vs. Luck: mischance hits ally
ℹtem[Life:] Growth surge, vines tether; vs. Death/Dreams: numbness, sleep−tug
ℹtem[Luck:] Odds flip; vs. Fate: harsher fixed outcome
ℹtem[Death/Dreams:] Whispers, chill; vs. Life: pain returns, rot
\end{description}

\section*{Magical Item Creation} ∈dex{magic!item creation}

Creating permanent magical items is a complex process.

\subsection*{Creation Requirements}

\begin{itemize}
ℹtem Knowledge: Understanding of the desired effect
ℹtem Materials: Appropriate components with magical properties
ℹtem Time: Significant investment of time and effort
ℹtem Skill: High level of magical and craft skills
ℹtem Facilities: Proper workspace with necessary tools
\end{itemize}

\subsection*{Creation Process}

\begin{enumerate}
ℹtem Design: Plan the item's properties and limitations
ℹtem Gathering: Acquire necessary materials and components
ℹtem Crafting: Physical creation of the item base
ℹtem Enchantment: Magical infusion of the desired properties
ℹtem Finishing: Final adjustments and testing
\end{enumerate}

\subsection*{Item Limitations}

\begin{itemize}
ℹtem Charges: Limited uses before needing recharge
ℹtem Attunement: Required bonding with the user
ℹtem Maintenance: Regular upkeep to preserve functionality
ℹtem Drawbacks: Negative side effects or requirements
\end{itemize}

\section*{Magic in Social Situations} ∈dex{magic!social}

Using magic in social contexts has special considerations.

\subsection*{Social Spellcasting}

\begin{itemize}
ℹtem Discretion: Avoiding detection while casting
ℹtem Consent: Ethical considerations of affecting others' minds
ℹtem Reactions: How different cultures view magical influence
ℹtem Laws: Legal restrictions on magical use in society
\end{itemize}

\subsection*{Social Backlash}

Magical social failures can cause:
\begin{itemize}
ℹtem Distrust: People becoming wary of the caster
ℹtem Resistance: Developing immunity or countermeasures
ℹtem Reputation: Becoming known as a manipulator
ℹtem Legal: Facing consequences from authorities
\end{itemize}

\section*{Learning and Improving Magic} ∈dex{magic!learning}

Magical ability grows through study and practice.

\subsection*{Skill Advancement}

\begin{itemize}
ℹtem Study: Researching magical theory and techniques
ℹtem Practice: Regular casting to improve control
ℹtem Experimentation: Trying new approaches and combinations
ℹtem Instruction: Learning from more experienced casters
\end{itemize}

\subsection*{Advanced Magical Development}

At higher levels, casters can:
\begin{itemize}
ℹtem Specialize: Focus on specific magical traditions
ℹtem Innovate: Create new spells or techniques
ℹtem Teach: Instruct others in magical arts
ℹtem Research: Discover lost or forbidden knowledge
\end{itemize}

\section*{Magical Safety and Ethics} ∈dex{magic!ethics}

Responsible magical practice involves understanding risks and consequences.

\subsection*{Safety Considerations}

\begin{itemize}
ℹtem Containment: Preventing unintended spread of effects
ℹtem Stability: Ensuring magical effects remain controlled
ℹtem Fail−safes: Planning for when magic goes wrong
ℹtem Recovery: Procedures for dealing with backlash
\end{itemize}

\subsection*{Ethical Guidelines}

\begin{itemize}
ℹtem Consent: Respecting others' autonomy regarding magic
ℹtem Transparency: Being honest about magical capabilities
ℹtem Restraint: Using magic judiciously and appropriately
ℹtem Responsibility: Accepting consequences of magical actions
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Magic Quick Reference, fonttitle=\bfseries]
\textbf{Casting (Freeform):}
\begin{itemize}
ℹtem Requires Talent: Spellcraft (6 XP)
ℹtem Weave & Cast: Two action effect using the Eight Elements
ℹtem Backlash: Any 1 rolled may cause narrative backlash
\end{itemize}

\textbf{Backlash Severity:}
\begin{itemize}
ℹtem On Partial/Miss: Pick 1−2 consequences flavored by Element
ℹtem Color consequences by Element (fire burns, fate twists, etc.)
\end{itemize}

\textbf{Rites System:}
\begin{itemize}
ℹtem Invoke: 1 action effect
ℹtem Obligation: Mark segments on timer
ℹtem Push It: +1 Obligation for +1 step effect
\end{itemize}

\textbf{Invoker Path:}
\begin{itemize}
ℹtem Symbols (4 XP each) grant ritual access
ℹtem Rituals: Significant Time, always +1 Obligation
ℹtem Crack the Seal: Instant cast (+2/+3 Obligation)
\end{itemize}

\textbf{Safety:} Every roll changes the story. Success without risk is rare.
\end{tcolorbox}

\section*{Practical Magic Examples} ∈dex{magic!examples}

\subsection*{Fire Cast, Partial}

You Weave flame to blind a squad (DV 3). Partial with two 1s. GM spends SB to Position −1 (flare blinds you too) and colors backlash as singed lashes; patrol is alerted (Exposure).

\subsection*{Runekeeper Push and Debt}

You Invoke Circle of Denial [WARD] and Push It to harden the ring. Mark +1 Obligation for the Rite plus +1 for the push. When a demon tests the ring, use [WARD] vs Cap; on its Hit, add +DV to its Leash.

\subsection*{Crack the Seal Under Fire}

You present Ikasha's Symbol and Crack the Seal to lay an instant shadow lane. Symbol → Compromised; mark +2 Obligation. GM immediately spends 1 SB to dim all lights (panic), then the lane forms. During downtime, you restore the Symbol (Arcana DV 3): a shaky hit leaves it Neglected until you perform the full rite of cleaning.

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Closing Thought, fonttitle=\bfseries]
Magic is a powerful tool but never a safe one. Every casting carries risks, and great power always demands great responsibility. Make bold choices−−−then let the consequences write the next chapter.
\end{tcolorbox}

\section*{Narrative−Heavy Magic Options} ∈dex{Narrative−Heavy Gameplay!Magic}

For groups that prefer strong narrative focus in magic use, consider these optional approaches:

\textbf{Intent−Driven Magic:} For minor magical effects that don't significantly alter the story, players can simply declare what they want to accomplish and describe how they do it, without rolling dice. The GM determines if the effect is reasonable and what complications might arise.

\textbf{Collaborative Backlash:} Instead of the GM unilaterally determining backlash, players can suggest thematic consequences that fit the fiction, with GM approval. This makes magic feel more collaborative and story−driven.

\textbf{Ritual as Story Beats:} Major magical workings can be treated as scene−defining moments where the group collaboratively describes what happens, with mechanical effects determined by the narrative impact rather than detailed rolls.

\textbf{Patron Relationships:} Focus on the roleplaying aspects of Patron relationships, treating Obligation as a measure of story tension and character development rather than just a mechanical track to be managed.

\textbf{Magic as Character Development:} Use magical experiences as opportunities for character growth and backstory development, allowing players to narrate how their characters learned new abilities through significant story moments.

\chapter{Attributes and Skills} łabel{ch:attributes}

Your character's capabilities are built on four core Attributes and specialized Skills. This chapter explains how they work together to define what your character can do and how they interact with the world, with clear examples and player−facing tips.

\section*{Core Attributes} ∈dex{Attributes}

Attributes represent your character's fundamental capabilities. Each is rated from 1 to 5, with higher numbers indicating greater proficiency.

\subsection*{Body}

Physical strength, endurance, coordination, and health.
\begin{itemize}
ℹtem Used for: Melee combat, athletics, endurance tests, physical labor
ℹtem Typical applications: Lifting, running, climbing, fighting, resisting physical harm
ℹtem Associated skills: Athletics, Brawl, Melee, Endurance
\end{itemize}

\textbf{Rating examples:}
\begin{description}
ℹtem[1:] Average person, some physical activity
ℹtem[2:] Fit individual, regular training
ℹtem[3:] Athlete or soldier, excellent condition
ℹtem[4:] Exceptional athlete, near−peak human
ℹtem[5:] Peak human capability, legendary strength
\end{description}

\subsection*{Wits}

Mental acuity, perception, quick thinking, and problem−solving.
\begin{itemize}
ℹtem Used for: Investigation, perception, tactics, quick decisions
ℹtem Typical applications: Spotting details, solving puzzles, planning, reacting quickly
ℹtem Associated skills: Perception, Investigation, Tactics, Lore
\end{itemize}

\textbf{Rating examples:}
\begin{description}
ℹtem[1:] Average awareness, sometimes misses things
ℹtem[2:] Observant, notices important details
ℹtem[3:] Sharp−minded, quick to spot patterns
ℹtem[4:] Exceptionally perceptive, rarely surprised
ℹtem[5:] Near−prescient awareness, sees connections others miss
\end{description}

\subsection*{Spirit}

Willpower, intuition, mental resilience, and connection to intangible forces.
\begin{itemize}
ℹtem Used for: Resisting mental effects, intuition, magical aptitude, determination
ℹtem Typical applications: Resisting fear, sensing danger, magical ability, enduring hardship
ℹtem Associated skills: Resolve, Intuition, Magic, Faith
\end{itemize}

\textbf{Rating examples:}
\begin{description}
ℹtem[1:] Average willpower, somewhat suggestible
ℹtem[2:] Strong−minded, resists ordinary pressure
ℹtem[3:] Very determined, hard to intimidate
ℹtem[4:] Exceptional will, inspires others
ℹtem[5:] Iron will, nearly unshakeable resolve
\end{description}

\subsection*{Presence}

Charisma, social influence, appearance, and force of personality.
\begin{itemize}
ℹtem Used for: Social interactions, leadership, persuasion, intimidation
ℹtem Typical applications: Negotiating, leading, charming, commanding attention
ℹtem Associated skills: Sway, Command, Performance, Deception
\end{itemize}

\textbf{Rating examples:}
\begin{description}
ℹtem[1:] Average presence, doesn't stand out
ℹtem[2:] Noticeable, makes an impression
ℹtem[3:] Charismatic, naturally influential
ℹtem[4:] Commanding presence, people listen
ℹtem[5:] Magnetic personality, can sway crowds
\end{description}

\section*{Skill System} ∈dex{Skills}

Skills represent specialized training and expertise. They combine with Attributes to form your dice pool for actions.

\subsection*{Skill Ratings}

\begin{center}
\begin{tabular}{|c|l|}
ℎline
\textbf{Rating} & \textbf{Description} \\
ℎline
0 & Untrained −−− No formal training \\
1 & Novice −−− Basic understanding \\
2 & Competent −−− Reliable skill level \\
3 & Professional −−− Expert capability \\
4 & Master −−− Renowned expertise \\
5 & Grand Master −−− Legendary skill \\
ℎline
\end{tabular}
\end{center}

\subsection*{Skill Categories}

⫕section*{Combat Skills}

\begin{itemize}
ℹtem Melee: Swords, axes, close−quarters weapons
ℹtem Ranged: Bows, crossbows, thrown weapons
ℹtem Brawl: Unarmed combat, grappling
ℹtem Tactics: Battlefield strategy, unit coordination
\end{itemize}

⫕section*{Physical Skills}

\begin{itemize}
ℹtem Athletics: Running, climbing, jumping
ℹtem Stealth: Moving unseen, hiding
ℹtem Endurance: Resisting fatigue, harsh conditions
ℹtem Craft: Building, repairing, creating
\end{itemize}

⫕section*{Social Skills}

\begin{itemize}
ℹtem Sway: Persuasion, negotiation, charm
ℹtem Command: Leadership, intimidation, authority
ℹtem Deception: Lying, bluffing, misdirection
ℹtem Performance: Entertainment, oration, acting
\end{itemize}

⫕section*{Knowledge Skills}

\begin{itemize}
ℹtem Lore: History, culture, general knowledge
ℹtem Investigation: Research, deduction, analysis
ℹtem Medicine: Healing, anatomy, treatment
ℹtem Nature: Wilderness, animals, plants
\end{itemize}

⫕section*{Specialized Skills}

\begin{itemize}
ℹtem Arcana: Magic, rituals, mystical knowledge
ℹtem Mechanics: Devices, engineering, construction
ℹtem Diplomacy: Formal negotiation, protocol
ℹtem Streetwise: Urban survival, criminal knowledge
\end{itemize}

\section*{Building Dice Pools} ∈dex{Dice Pool}

Your dice pool for any action is: Attribute + Skill.

\subsection*{Choosing the Right Combination}

The same action can often be approached with different Attribute/Skill combinations:
\begin{itemize}
ℹtem Climbing a wall:
\begin{itemize}
ℹtem Body + Athletics (physical strength)
ℹtem Wits + Athletics (finding the best route)
ℹtem Spirit + Athletics (sheer determination)
\end{itemize}
ℹtem Persuading a guard:
\begin{itemize}
ℹtem Presence + Sway (charm and personality)
ℹtem Wits + Sway (logical arguments)
ℹtem Spirit + Sway (force of conviction)
\end{itemize}
ℹtem Investigating a crime scene:
\begin{itemize}
ℹtem Wits + Investigation (careful observation)
ℹtem Spirit + Investigation (intuitive leaps)
ℹtem Presence + Investigation (getting people to talk)
\end{itemize}
\end{itemize}

\subsection*{Creative Combinations}

With GM approval, you can justify unusual combinations:
\begin{itemize}
ℹtem Body + Lore for recalling physical techniques
ℹtem Presence + Medicine for comforting patients
ℹtem Spirit + Craft for inspired artistic creation
\end{itemize}

\textbf{Example:} A ranger scales an ice wall using Wits + Athletics to route−find, then switches to Body + Athletics to muscle over the lip. The fiction guides the mechanics.

\section*{Skill Advancement} ∈dex{Skills!advancement}

Improving skills requires experience points and training time.

\subsection*{XP Costs}

\begin{center}
\begin{tabular}{|l|r|}
ℎline
\textbf{Improvement} & \textbf{XP Cost} \\
ℎline
0→1 & 2 XP \\
1→2 & 4 XP \\
2→3 & 6 XP \\
3→4 & 8 XP \\
4→5 & 10 XP \\
ℎline
\end{tabular}
\end{center}

\subsection*{Training Time}

\begin{itemize}
ℹtem 0 → 1: 1 day of practice
ℹtem 1 → 2: 3 days of training
ℹtem 2 → 3: 1 week of intensive study
ℹtem 3 → 4: 2 weeks of master training
ℹtem 4 → 5: 1 month of dedicated practice
\end{itemize}

\subsection*{Attribute Limits}

You cannot have a skill rating higher than its primary Attribute. To increase a skill beyond your Attribute, you must first improve the Attribute.

\section*{Synergy Between Skills} ∈dex{Skills!synergy}

Some skills work particularly well together, providing bonuses when used in combination. Synergies are situational and require fictional justification.

\subsection*{Combat Synergies}

\begin{itemize}
ℹtem Tactics + Command: +1 die when leading groups in combat
ℹtem Melee + Athletics: +1 die on movement−based attacks
ℹtem Ranged + Perception: +1 die on aimed shots
\end{itemize}

\subsection*{Social Synergies}

\begin{itemize}
ℹtem Sway + Lore: +1 die when using knowledge in persuasion
ℹtem Deception + Performance: +1 die on sustained deceptions
ℹtem Command + Presence: +1 die on leadership actions
\end{itemize}

\subsection*{Exploration Synergies}

\begin{itemize}
ℹtem Investigation + Perception: +1 die on detailed searches
ℹtem Nature + Survival: +1 die on wilderness navigation
ℹtem Mechanics + Craft: +1 die on complex repairs
\end{itemize}

\section*{Using Skills in Play} ∈dex{Skills!usage}

\subsection*{When to Roll}

Skills are used when:
\begin{itemize}
ℹtem The outcome is uncertain
ℹtem There are meaningful consequences for failure
ℹtem The action is significant to the story
\end{itemize}

\subsection*{Difficulty Values by Skill Level}

\begin{center}
\begin{tabular}{|l|c|c|}
ℎline
\textbf{Skill Level} & \textbf{Routine Task} & \textbf{Challenging Task} \\
ℎline
0 & DV 2 & DV 4 \\
1 & DV 1 & DV 3 \\
2 & Automatic & DV 2 \\
3 & Automatic & DV 1 \\
4+ & Automatic & Automatic \\
ℎline
\end{tabular}
\end{center}

\textbf{Reading the Table:} A Professional (3) auto−succeeds on routine tasks; challenge them with interesting stakes or higher DVs.

\subsection*{Group Skill Use}

When multiple characters use the same skill:
\begin{itemize}
ℹtem Assistance: One character leads, others provide +1 die each (max +3)
ℹtem Cooperation: Multiple characters attempt the same task separately
ℹtem Complementary: Different skills used together for a complex task
\end{itemize}

\section*{Skill Challenges} ∈dex{Skills!challenges}

Complex tasks may require multiple skill uses or extended effort.

\subsection*{Extended Tests}

For tasks taking significant time:
\begin{itemize}
ℹtem Set a timer with 4−−8 segments
ℹtem Each successful skill use fills segments
ℹtem Complications may add segments or create setbacks
\end{itemize}

\subsection*{Complex Challenges}

Tasks requiring multiple skills:
\begin{itemize}
ℹtem Different characters use different skills
ℹtem Successes contribute to overall progress
ℹtem Failure in one area may complicate others
\end{itemize}

\textbf{Example (Complex Heist):} Stealth to enter, Mechanics to bypass locks, Investigation to locate the vault, Deception to mislead guards. Each success advances the Heist Timer ; SB creates new heat.

\section*{Skill−Based Character Archetypes} ∈dex{Character Archetypes}

\subsection*{The Warrior}

\begin{itemize}
ℹtem Primary: Body + Melee/Ranged
ℹtem Secondary: Spirit + Endurance, Wits + Tactics
ℹtem Key skills: Athletics, Brawl, Command
ℹtem Playstyle: Direct confrontation, physical solutions
\end{itemize}

\subsection*{The Expert}

\begin{itemize}
ℹtem Primary: Wits + Lore/Investigation
ℹtem Secondary: Presence + Sway, Spirit + Resolve
ℹtem Key skills: Mechanics, Medicine, Perception
ℹtem Playstyle: Problem−solving, information gathering
\end{itemize}

\subsection*{The Face}

\begin{itemize}
ℹtem Primary: Presence + Sway/Deception
ℹtem Secondary: Wits + Investigation, Spirit + Performance
ℹtem Key skills: Command, Diplomacy, Streetwise
ℹtem Playstyle: Social manipulation, negotiation
\end{itemize}

\subsection*{The Specialist}

\begin{itemize}
ℹtem Primary: Varies by specialty
ℹtem Secondary: Supporting skills for the specialty
ℹtem Key skills: Craft, Arcana, Nature, etc.
ℹtem Playstyle: Technical expertise, unique capabilities
\end{itemize}

\section*{Improving Your Capabilities} ∈dex{Character Development}

\subsection*{Balanced Development}

\begin{itemize}
ℹtem Improve both Attributes and Skills together
ℹtem Develop complementary skill sets
ℹtem Consider how skills work in combination
ℹtem Plan for both immediate needs and long−term growth
\end{itemize}

\subsection*{Specialized Focus}

\begin{itemize}
ℹtem Maximize one Attribute and related skills
ℹtem Develop deep expertise in one area
ℹtem Become the go−to character for specific challenges
ℹtem Risk being less effective outside your specialty
\end{itemize}

\subsection*{Versatile Approach}

\begin{itemize}
ℹtem Moderate investment in multiple areas
ℹtem Ability to handle diverse situations
ℹtem Less peak capability but more adaptability
ℹtem Good for supporting other characters
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Attributes and Skills Quick Reference, fonttitle=\bfseries]
\textbf{Attributes (1−−5):}
\begin{itemize}
ℹtem Body: Physical capability
ℹtem Wits: Mental acuity
ℹtem Spirit: Willpower
ℹtem Presence: Social influence
\end{itemize}

\textbf{Skill Levels:}
\begin{itemize}
ℹtem 0: Untrained | 1: Novice | 2: Competent
ℹtem 3: Professional | 4: Master | 5: Grand Master
\end{itemize}

\textbf{Dice Pool:} Attribute + Skill d10s\\
\textbf{Improvement:} New level × 2 XP (skills)\\
\textbf{Specialization:} +1 die in specific area at level 3+\\
\textbf{Synergy:} Complementary skills give +1 die
\end{tcolorbox}

\section*{Practical Examples} ∈dex{Skills!examples}

\subsection*{Combat Example}

A warrior (Body 4, Melee 3) attacks:
\begin{itemize}
ℹtem Dice pool: 4 + 3 = 7d10
ℹtem Needs 6+ on each die for successes
ℹtem DV set by opponent's defense (typically 2−−3)
\end{itemize}

\subsection*{Social Example}

A diplomat (Presence 3, Sway 2) negotiates:
\begin{itemize}
ℹtem Dice pool: 3 + 2 = 5d10
ℹtem Position: Controlled (opponent is skeptical)
ℹtem Stakes: Success gets cooperation, failure creates suspicion
\end{itemize}

\subsection*{Exploration Example}

A scout (Wits 3, Perception 2) searches for tracks:
\begin{itemize}
ℹtem Dice pool: 3 + 2 = 5d10
ℹtem DV 2 for fresh tracks, DV 3 for old tracks
ℹtem Success finds trail, partial finds clues, miss misses important signs
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Final Note, fonttitle=\bfseries]
Your Attributes and Skills define not just what you can do, but how you approach challenges. Choose combinations that reflect your character's personality and style, and let the fiction lead your mechanical choices.
\end{tcolorbox}

\section*{Narrative−Heavy Skill Options} ∈dex{Narrative−Heavy Gameplay!Skills}

For groups that prefer strong narrative focus in skill use, consider these optional approaches:

\textbf{Intent−Driven Skills:} For routine tasks that don't significantly impact the story, players can simply declare what they want to accomplish without rolling dice. The GM determines if the action succeeds based on the character's capabilities and the fiction.

\textbf{Descriptive Assistance Bonuses:} Players can provide vivid, helpful descriptions of how they're using their skills to assist allies, granting a +1 die bonus to the primary actor's roll without spending Boons.

\textbf{Skill as Character Development:} Use skill challenges as opportunities for character growth and backstory development, allowing players to narrate how their characters learned new techniques through significant story moments.

\textbf{Collaborative Difficulty Setting:} Instead of the GM unilaterally setting DVs, players can suggest reasonable difficulty levels based on their understanding of the task, with GM approval.

\textbf{Narrative Skill Synergies:} Focus on how skills work together in the story rather than mechanical bonuses. A well−described combination of skills might grant advantage on Position or Effect without requiring specific synergy rules.

\chapter{Experience Paths and Character Building} łabel{ch:paths}

How you spend your Experience Points (XP) defines not only your character's capabilities−−−but also their role in the world. This chapter explores different advancement philosophies and provides practical, legal starting builds that fit the campaign's creation rules.

\section*{Three Advancement Paths} ∈dex{Character Development Paths}

There are three broad approaches to character development, each representing a different philosophy of growth:

\begin{description}
ℹtem[Personal Path] Invest in personal mastery and self−improvement
ℹtem[Balanced Path] Mix personal growth with resources and influence
ℹtem[Influencer Path] Focus on networks, assets, and strategic power
\end{description}

\section*{Path 1: Personal Development} ∈dex{Character Development Paths!Personal}

The Personal Path focuses on individual capability through attributes and skills.

\subsection*{Typical Investment}

\begin{itemize}
ℹtem 70−−90% Personal improvement
ℹtem 0−−10% Resources and assets
ℹtem 0−−20% Special abilities
\end{itemize}

\subsection*{Strengths}

\begin{itemize}
ℹtem Reliable in direct challenges and combat
ℹtem Minimal upkeep or management required
ℹtem Resilient to loss of external resources
ℹtem Consistent performance in spotlight moments
\end{itemize}

\subsection*{Weaknesses}

\begin{itemize}
ℹtem Limited influence in social or strategic scenes
ℹtem May struggle with problems requiring networks
ℹtem Less capable in logistics or large−scale operations
ℹtem Dependent on personal presence for all solutions
\end{itemize}

\subsection*{Build Example: The Duelist (Legal Start)}

Total XP: 30 (36with +4 from Bonds/Complications; see §6.5)

\begin{itemize}
ℹtem Attributes: Body 3, Wits 2, Spirit 1, Presence 1
\begin{itemize}
ℹtem Costs (Attributes cost new rating ×3 each step): Body 1→2 (6), 2→3 (9) = 15; Wits 1→2 (6) = 6; Spirit/Presence remain 1 = 0. Subtotal: 21 XP
\end{itemize}
ℹtem Skills: Melee 2, Athletics 1
\begin{itemize}
ℹtem Costs (Skills cost new level ×2 each step): Melee 0→1 (2), 1→2 (4) = 6; Athletics 0→1 = 2. Subtotal: 8 XP
\end{itemize}
ℹtem Totals: 21 + 8 = 29 XP. Bank 1
```latex
 XP.
ℹtem With +4 XP (Bonds/Complications): add Perception 0→1 (2) and spend banked 1 XP on Stealth 0→1 (2), or instead take Perception 0→1 (2) and Sway 0→1 (2) for broader utility.
ℹtem Cap: 36XP.
\end{itemize}

\section*{Path 2: Balanced Approach} ∈dex{Character Development Paths!Balanced}

The Balanced Path mixes personal capability with strategic resources.

\subsection*{Typical Investment}

\begin{itemize}
ℹtem 50−−65% Personal improvement
ℹtem 15−−░ Resources and assets
ℹtem 15−−░ Special abilities
\end{itemize}

\subsection*{Strengths}

\begin{itemize}
ℹtem Adaptable to diverse situations
ℹtem Can handle both direct and indirect challenges
ℹtem Good supporting role for the group
ℹtem Moderate risk profile
\end{itemize}

\subsection*{Weaknesses}

\begin{itemize}
ℹtem Not exceptional in any single area
ℹtem Requires management of resources
ℹtem Moderate upkeep demands
ℹtem Can be outshone by specialists
\end{itemize}

\subsection*{Build Example: The Scout (Legal Start)}

Total XP: 30 (36with +4 from Bonds/Complications)

\begin{itemize}
ℹtem Attributes: Wits 2, Body 2, Spirit 1, Presence 1
\begin{itemize}
ℹtem Costs: Wits 1→2 (6), Body 1→2 (6) = 12 XP
\end{itemize}
ℹtem Skills: Survival 2, Perception 1, Stealth 1
\begin{itemize}
ℹtem Costs: Survival 0→1 (2), 1→2 (4) = 6; Perception 0→1 2; Stealth 0→1 2. Subtotal: 10 XP
\end{itemize}
ℹtem Resources: Minor equipment cache (camp gear, maps, signal kit) = 4 XP
ℹtem Special Abilities: Wilderness Lore (broad travel benefits) = 4 XP
ℹtem Totals: 12 + 10 + 4 + 4 = 32 XP.
ℹtem With +4 XP: add Perception 1→2 (+4) or take a trained hawk companion (Minor Resource, 4 XP).
\end{itemize}

\section*{Path 3: Influencer Focus} ∈dex{Character Development Paths!Influencer}

The Influencer Path prioritizes networks, assets, and strategic power.

\subsection*{Typical Investment}

\begin{itemize}
ℹtem 25−−40% Personal improvement
ℹtem 35−−55% Resources and assets
ℹtem 20−−40% Special abilities
\end{itemize}

\subsection*{Strengths}

\begin{itemize}
ℹtem Strong strategic and social influence
ℹtem Can solve problems indirectly
ℹtem Excellent at planning and preparation
ℹtem Creates opportunities for the whole group
\end{itemize}

\subsection*{Weaknesses}

\begin{itemize}
ℹtem Personally vulnerable in direct confrontations
ℹtem High maintenance requirements
ℹtem Complications can cascade through networks
ℹtem Dependent on external factors
\end{itemize}

\subsection*{Build Example: The Merchant (Legal Start)}

Total XP: 30 (36with +4 from Bonds/Complications)

\begin{itemize}
ℹtem Attributes: Presence 2, Wits 2, Spirit 1, Body 1
\begin{itemize}
ℹtem Costs: Presence 1→2 (6), Wits 1→2 (6) = 12 XP
\end{itemize}
ℹtem Skills: Sway 2, Deception 1, Lore 1
\begin{itemize}
ℹtem Costs: Sway 0→1 (2), 1→2 (4) = 6; Deception 0→1 2; Lore 0→1 2. Subtotal: 10 XP
\end{itemize}
ℹtem Resources: Standard trading office (staffed storefront, ledgers, storage) = 8 XP
ℹtem Totals: 12 + 10 + 8 = 32 XP.
ℹtem With +4 XP: add Negotiation Mastery (4 XP general ability) or expand to a second Minor merchant route (4 XP).
\end{itemize}

\section*{Starting Character Guidelines} ∈dex{Character Creation}

\subsection*{Base XP Allocation}

\begin{itemize}
ℹtem Standard Starting XP: 30 points
ℹtem Bonds and Complications: You may take up to two total from any mix of meaningful Bonds (up to 2, +2 XP each) and significant Complications (up to 2, +2 XP each), granting maximum +4 XP.
ℹtem Maximum Starting XP: 36points
ℹtem Complication Effect: Each unresolved starting Complication adds +1 banked SB to early scenes until cleared.
\end{itemize}

\subsection*{Recommended Starting Ranges}

\begin{center}
\begin{tabular}{|l|r|}
ℎline
\textbf{Category} & \textbf{Recommended XP} \\
ℎline
Primary Attribute & 9−−12 XP (rating 3−−4) \\
Secondary Attributes & 0−−9 XP each (rating 1−−3) \\
Key Skills & 4−−6 XP each (rating 2−−3) \\
Supporting Skills & 2−−4 XP each (rating 1−−2) \\
Resources & 0−−8 XP total \\
Special Abilities & 0−−8 XP total \\
ℎline
\end{tabular}
\end{center}

\subsection*{Cost Reminders:}

\begin{itemize}
ℹtem Attributes: Each step costs new rating ×3 XP (e.g., 1→2 costs 6; 2→3 costs 9).
ℹtem Skills: Each step costs new level ×2 XP (e.g., 0→1 costs 2; 1→2 costs 4).
ℹtem Resources: Minor 4 XP; Standard 8 XP; Major 12 XP.
ℹtem Special Abilities: Minor Edge 2 XP; Major Edge 4 XP; Prestige 6+ XP.
\end{itemize}

\section*{Progression Planning} ∈dex{Character Development!Planning}

\subsection*{Early Game (0−−40 XP)}

Focus on establishing core capabilities:
\begin{itemize}
ℹtem Reach attribute rating 3 in your primary area
ℹtem Develop 2−−3 key skills to rating 2−−3
ℹtem Acquire basic resources or one special ability
ℹtem Establish your character's niche in the group
\end{itemize}

\subsection*{Mid Game (41−−90 XP)}

Expand and specialize:
\begin{itemize}
ℹtem Increase primary attribute to 4
ℹtem Specialize key skills to rating 3−−4
ℹtem Develop supporting capabilities
ℹtem Build strategic resources or networks
ℹtem Acquire signature special abilities
\end{itemize}

\subsection*{Late Game (91−−150 XP)}

Master your chosen path:
\begin{itemize}
ℹtem Achieve peak attributes (rating 4−−5)
ℹtem Master key skills (rating 4−−5)
ℹtem Build substantial influence or unique capabilities
ℹtem Develop advanced special abilities
ℹtem Consider legacy projects or organizations
\end{itemize}

\section*{Path Combination Strategies} ∈dex{Character Development!Combination}

Many players mix elements from different paths:

\subsection*{Combat Specialist with Resources}

\begin{itemize}
ℹtem Strong personal combat capabilities
ℹtem Moderate resource investment for support
ℹtem Good for frontline fighters who need logistical support
ℹtem Example: Warrior with a fortified base and loyal troops
\end{itemize}

\subsection*{Social Character with Personal Skills}

\begin{itemize}
ℹtem Excellent social capabilities
ℹtem Solid personal skills for self−defense
ℹtem Good for diplomats who operate independently
ℹtem Example: Ambassador with combat training and persuasion skills
\end{itemize}

\subsection*{Technical Expert with Networks}

\begin{itemize}
ℹtem Deep technical or magical expertise
ℹtem Network of contacts and resources
ℹtem Good for specialists who need support systems
ℹtem Example: Master crafter with supplier network and apprentices
\end{itemize}

\section*{Resource Management} ∈dex{Resource Management}

Each path requires different management approaches:

\subsection*{Personal Path Management}

\begin{itemize}
ℹtem Minimal upkeep requirements
ℹtem Focus on equipment maintenance
ℹtem Occasional skill practice or training
ℹtem Low complexity, high reliability
\end{itemize}

\subsection*{Balanced Path Management}

\begin{itemize}
ℹtem Moderate upkeep for resources
ℹtem Relationship maintenance with contacts
ℹtem Skill development alongside resource management
ℹtem Balanced time investment
\end{itemize}

\subsection*{Influencer Path Management}

\begin{itemize}
ℹtem Significant upkeep demands
ℹtem Network maintenance and expansion
ℹtem Resource allocation and development
ℹtem Strategic planning and opportunity management
\end{itemize}

\section*{Risk Assessment} ∈dex{Character Development!Risk}

Each path carries different risks:

\subsection*{Personal Path Risks}

\begin{itemize}
ℹtem Over−specialization in one area
ℹtem Vulnerability to problems outside specialty
ℹtem Limited growth options later in game
ℹtem May become predictable in approach
\end{itemize}

\subsection*{Balanced Path Risks}

\begin{itemize}
ℹtem Jack−of−all−trades, master of none
ℹtem Spread too thin across capabilities
ℹtem Moderate risks in multiple areas
ℹtem May lack standout capabilities
\end{itemize}

\subsection*{Influencer Path Risks}

\begin{itemize}
ℹtem Network vulnerability to attacks
ℹtem High maintenance requirements
ℹtem Cascade failure potential
ℹtem Personal safety concerns
\end{itemize}

\section*{Building for Group Synergy} ∈dex{Group Dynamics}

Consider how your path complements other party members:

\subsection*{Complementary Paths}

\begin{itemize}
ℹtem Personal path characters provide reliable combat capability
ℹtem Balanced path characters handle diverse challenges
ℹtem Influencer path characters create opportunities and resources
ℹtem Mixed groups cover all bases effectively
\end{itemize}

\subsection*{Redundant Paths}

\begin{itemize}
ℹtem Multiple personal path characters may overlap in combat
ℹtem Multiple influencer path characters may compete for resources
ℹtem Consider diversifying within similar paths
ℹtem Example: Different combat specialties or resource types
\end{itemize}

\section*{Adapting Your Path} ∈dex{Character Development!Adaptation}

Your chosen path isn't permanent−−−you can shift focus as the game progresses:

\subsection*{Early Shift (0−−40 XP)}

\begin{itemize}
ℹtem Easy to change direction
ℹtem Minimal sunk cost in any approach
ℹtem Good time to experiment with different styles
ℹtem Can respond to group needs or story developments
\end{itemize}

\subsection*{Mid Game Shift (41−−90 XP)}

\begin{itemize}
ℹtem Requires more deliberate planning
ℹtem Some capabilities may need to be maintained
ℹtem Can fill emerging gaps in group capability
ℹtem May require temporary performance dip during transition
\end{itemize}

\subsection*{Late Game Shift (91+ XP)}

\begin{itemize}
ℹtem Significant investment in current path
ℹtem Major shift requires substantial XP investment
ℹtem Consider adding complementary capabilities rather than replacing
ℹtem May be better to develop existing strengths further
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=XP Path Quick Reference, fonttitle=\bfseries]
\textbf{Personal Path (70−−90% self):}
\begin{itemize}
ℹtem Reliable individual performance
ℹtem Low upkeep, high consistency
ℹtem Best for combat and specialist roles
\end{itemize}

\textbf{Balanced Path (50−−65% self):}
\begin{itemize}
ℹtem Good all−around capability
ℹtem Moderate risk and upkeep
ℹtem Flexible supporting role
\end{itemize}

\textbf{Influencer Path (25−−40% self):}
\begin{itemize}
ℹtem Strategic power and influence
ℹtem High upkeep, high reward
ℹtem Creates opportunities for group
\end{itemize}

\textbf{Starting XP:} 30 base + up to +4 from Bonds/Complications (max start 34).
\end{tcolorbox}

\section*{Practical Building Examples (Narrative Roles, Legal Starts)} ∈dex{Character Creation!Examples}

\subsection*{Example 1: The Guardian}

Path: Personal ℎfill Total: 32 XP

\begin{itemize}
ℹtem Attributes: Body 3 (15), Wits 2 (6) = 21 XP
ℹtem Skills: Melee 2 (6), Athletics 1 (2) = 8 XP
ℹtem Bank: 1 XP
ℹtem Role at table: Frontline protection, reliable duel pressure. With +4 XP, add Combat Reflexes (2 XP talent) and Shield Mastery (4 XP talent) using banked 1 + 4 = 5 XP, with 1 XP remaining.
\end{itemize}

\subsection*{Example 2: The Explorer}

Path: Balanced ℎfill Total: 32 XP

\begin{itemize}
ℹtem Attributes: Wits 2 (6), Body 2 (6) = 12 XP
ℹtem Skills: Survival 2 (6), Perception 1 (2), Stealth 1 (2) = 10 XP
ℹtem Resources: Minor mapping kit & route notes = 4 XP
ℹtem Ability: Trail Sense = 4 XP
ℹtem Totals: 32 XP. With +4 XP, raise Perception 1→2 (+4) or add a trained beast (Minor Resource, 4).
\end{itemize}

\subsection*{Example 3: The Schemer}

Path: Influencer ℎfill Total: 32 XP

\begin{itemize}
ℹtem Attributes: Presence 2 (6), Wits 2 (6) = 12 XP
ℹtem Skills: Sway 2 (6), Deception 1 (2), Lore 1 (2) = 10 XP
ℹtem Resources: Standard safehouse & message drops = 8 XP
ℹtem Totals: 32 XP. With +4 XP, take Network Builder (4 XP talent) or add Minor informant ring (4 XP).
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Reminder, fonttitle=\bfseries]
All builds above assume baseline Attributes at 1 and Skills at 0 before spending. Attribute and Skill advances are cumulative by step (see costs in §6.5).

Remember: Your chosen path should reflect both your character concept and your preferred play style. There's no single "correct" path−−−only what works for you and your group.
\end{tcolorbox}

\section*{Narrative−Heavy Character Building Options} ∈dex{Narrative−Heavy Gameplay!Character Building}

For groups that prefer strong narrative focus in character building, consider these optional approaches:

\textbf{Story−Driven Milestones:} Instead of tracking XP numerically, the GM can award advancement when characters reach significant story milestones. "You've trained with the master for months−−−you've improved your skill."

\textbf{Experience Through Reflection:} Players can spend downtime scenes reflecting on past experiences to earn XP. A meaningful flashback or character moment can justify growth without tracking specific points.

\textbf{Collaborative Advancement:} The group can discuss and agree on advancement choices, ensuring everyone's growth supports the overall story direction.

\textbf{Narrative Justification Focus:} When spending XP, players should explain how their character gained this capability through in−game experiences, creating richer backstory and continuity.

\textbf{Path as Theme:} Focus on the narrative themes of your chosen path rather than strict XP allocations. A Personal Path character might emphasize their journey of self−mastery, while an Influencer Path character focuses on their growing web of relationships and influence.

\chapter{Talents and Special Abilities} łabel{ch:talents}

Talents are the building blocks of character specialization. They represent learned techniques, supernatural gifts, or cultural inheritances. Each Talent costs XP, and their costs are tied to impact.

\section*{Understanding Talents} ∈dex{Talents}

Talents are purchased with Experience Points (XP) and provide special capabilities:
\begin{itemize}
ℹtem They go beyond simple skill bonuses
ℹtem They often have specific activation conditions
ℹtem They may provide narrative permissions (you can try things others cannot)
ℹtem They can define your character's unique identity
\end{itemize}

\subsection*{Talent Costs}

\begin{center}
\begin{tabular}{|l|r|l|}
ℎline
\textbf{Type} & \textbf{Cost} & \textbf{Examples} \\
ℎline
Minor Edge & 6 XP & Spellcraft, +1 situational bonus \\
Major Edge & 4 XP & Patron's Symbol, strong summon upgrade \\
Prestige & 6+ XP & Campaign−defining effects \\
ℎline
\end{tabular}
\end{center}

\subsection*{Activation Types}

\begin{description}
ℹtem[Passive:] Always on; no action
ℹtem[Active:] Requires an action or scene focus
ℹtem[Reactive:] Triggers on a condition
\end{description}

\subsection*{Limits and Economy}

Unless a talent says otherwise:
\begin{itemize}
ℹtem Per Scene uses refresh at scene end
ℹtem Per Session uses refresh after downtime
ℹtem Some talents allow you to spend Boons to push effects
\end{itemize}

\section*{Talent Categories} ∈dex{Talents!categories}

\subsection*{Minor Edge Talents}

Basic abilities available to any character:
\begin{itemize}
ℹtem Cost: 2 XP
ℹtem Examples: Spellcraft, Familiar Bond, basic magical abilities
ℹtem Best for: Essential capabilities and access requirements
\end{itemize}

\subsection*{Major Edge Talents}

Significant abilities with moderate requirements:
\begin{itemize}
ℹtem Cost: 4 XP
ℹtem Examples: Patron's Symbol, Codex, significant summon upgrades
ℹtem Best for: Core specialization and magical access
\end{itemize}

\subsection*{Prestige Talents}

Powerful abilities unlocked through mastery or story events:
\begin{itemize}
ℹtem Cost: 6+ XP
ℹtem Examples: Breaking fundamental limits, forbidden summons, rewriting obligations
ℹtem Best for: Campaign−shaping capabilities
\end{itemize}

\section*{Magic Access Talents} ∈dex{Talents!magic access}

\subsection*{Spellcraft}

Cost: 6 XP

Grants access to Weave & Cast freeform spellcasting using the Eight Elements. Without this, characters cannot freeform cast.

\subsection*{Familiar}

Cost: 2 XP

Required to access Patron features such as Patron's Gift. Binds a Thiasos.

\subsection*{Codex}

Cost: 4 XP

Required to fully join a Patron's service as a Runekeeper. Grants access to that Patron's Rites and Obligation system.

\subsection*{Patron's Symbol}

Cost: 4 XP

Minor Asset. Allows an Invoker to access a Patron's Rites via ritual precision. Each Patron requires its own Symbol.

\section*{Patron's Gift (Imbuement)} ∈dex{Talents!Patron's Gift}

Cost: Free (requires Thiasos)

Activation: 1 Action once per scene

Duration: Scene

Range: Touch

Effect: Imbue one item with temporary magical power related to your Patron's domain. The item functions as a magical weapon (+1 Melee) and specialized tool (+1 thematic Skill) for the scene.

Push It: The item's power persists for one additional scene but marks +1 Obligation.

\section*{Selecting Talents} ∈dex{Talents!selection}

\subsection*{Consider Your Magical Path}

Choose talents that reinforce your character's magical approach:
\begin{itemize}
ℹtem Caster: Freeform spellcasting talents, elemental control
ℹtem Runekeeper: Rites access, Obligation management, Patron specialization
ℹtem Invoker: Ritual efficiency, Symbol maintenance, invocation speed
ℹtem Specialist: Unique talents matching your specific focus
\end{itemize}

\subsection*{Balance Access and Power}

Consider both access requirements and power talents:
\begin{itemize}
ℹtem Access: Essential prerequisites (Spellcraft, Familiar)
ℹtem Power: Combat enhancements, magical amplifications
ℹtem Utility: Support abilities, resource management
\end{itemize}

\subsection*{Think About Investment}

Consider how much XP each talent represents:
\begin{itemize}
ℹtem Minor (2 XP): Essential access, small narrative tricks
ℹtem Major (4 XP): Strong upgrades, permanent effects in niche
ℹtem Prestige (6+ XP): Campaign−defining, fundamental limits broken
\end{itemize}

\section*{Talent Building Strategies} ∈dex{Talents!building strategies}

\subsection*{The Specialist}

Focus on talents supporting one primary magical path:
\begin{itemize}
ℹtem Choose talents that synergize with each other
ℹtem Develop a clear specialization identity
ℹtem Become the go−to character for specific magical challenges
ℹtem Risk: May be less effective outside specialty
\end{itemize}

\subsection*{The Generalist}

Spread talents across multiple magical approaches:
\begin{itemize}
ℹtem Cover different types of magical challenges
ℹtem Provide support to other party members
ℹtem Adapt to diverse situations
ℹtem Risk: Less peak capability in any area, increased bookkeeping
\end{itemize}

\subsection*{The Foundation Builder}

Focus on essential access talents first:
\begin{itemize}
ℹtem Prioritize access requirements (Spellcraft, Familiar)
ℹtem Build toward major capabilities
ℹtem Establish core identity before specialization
ℹtem Risk: May lack immediate power payoff
\end{itemize}

\section*{Talent Examples} ∈dex{Talents!examples}

\subsection*{Magic Access Talents}

\begin{description}
ℹtem[Spellcraft (6 XP)] −−− Access to Weave & Cast freeform spellcasting using the Eight Elements.
ℹtem[Familiar (2 XP)] −−− Required for Patron's Gift and other Patron features.
ℹtem[Codex (4 XP)] −−− Full access to a Patron's Rites and Obligation system.
ℹtem[Patron's Symbol (4 XP)] −−− Ritual access to a Patron's Rites via invocation.
\end{description}

\subsection*{Combat Talents}

\begin{description}
ℹtem[Second Wind (2 XP, Active)] −−− Once per scene, clear 1 Fatigue when you take a moment to catch your breath.
ℹtem[Combat Reflexes (2 XP, Reactive)] −−− +1 dv on defense when surprised or flanked.
ℹtem[Precise Strike (2 XP, Active)] −−− Once per scene, ignore armor on one attack if you had Dominant or Controlled position.
ℹtem[Weapon Mastery (4 XP, Passive)] −−− Choose a weapon type; +1 die when using it.
\end{description}

\subsection*{Social Talents}

\begin{description}
ℹtem[Silver Tongue (2 XP, Passive)] −−− +1 die on persuasion attempts.
ℹtem[Read Emotions (2 XP, Active)] −−− Once per scene, automatically detect surface emotions in a social exchange.
ℹtem[Command Presence (4 XP, Passive)] −−− +1 die on leadership and intimidation rolls.
ℹtem[Network Builder (4 XP, Passive)] −−− Gain a minor contact in each new settlement visited.
\end{description}

\subsection*{Exploration Talents}

\begin{description}
ℹtem[Keen Senses (2 XP, Passive)] −−− +1 die on perception checks to spot danger or hidden details.
ℹtem[Wilderness Lore (2 XP, Passive)] −−− Automatically find food and water in hospitable biomes.
ℹtem[Trackless Step (2 XP, Active)] −−− Leave no trail for the rest of the scene.
ℹtem[Urban Navigation (2 XP, Passive)] −−− Never get lost in cities.
\end{description}

\section*{Advanced Talent Examples} ∈dex{Talents!advanced examples}

\subsection*{Casting Mastery}

\begin{description}
ℹtem[Spell Shaping (4 XP; Req: Spellcraft)] −−− Modify spell factors (range/scale/targeting) by one step when you Weave.
ℹtem[Elemental Mastery (6 XP; Req: Arcana 3)] −−− Reduce backlash severity by one step when casting spells of your chosen element.
ℹtem[Arcane Dominance (6 XP; Req: Spirit 4, Arcana 4)] −−− Overpower weaker magical effects automatically when you contest them.
\end{description}

\subsection*{Ritual Expertise}

\begin{description}
ℹtem[Ritual Mastery (4 XP; Req: Familiar)] −−− Perform rituals with reduced risk: the GM spends 1 fewer SB on ritual backlash.
ℹtem[Efficient Invocation (4 XP; Req: Patron's Symbol)] −−− Reduce ritual casting time by one step (minimum 1 Player Turn).
ℹtem[Crack Specialist (6 XP; Req: 3 Patron Symbols)] −−− Reduce Crack the Seal Obligation cost by 1 (minimum +1).
ℹtem[Dual Covenant (6 XP):] Maintain two active summons.
\end{description}

\subsection*{Prestige Abilities}

\begin{description}
ℹtem[Forbidden Knowledge (6 XP; Req: Tier II)] −−− Access to one forbidden summon or dangerous rite.
ℹtem[Obligation Master (8 XP; Req: Tier III, Codex)] −−− Reduce all Obligation segment costs by 1 (minimum 1).
ℹtem[Backlash Immunity (10 XP; Req: Tier IV, Spirit 5)] −−− Ignore minor backlash entirely on casting rolls.
ℹtem[Triad Bond (8 XP):] Maintain three active summons.
\end{description}

\section*{Talent Synergies} ∈dex{Talents!synergies}

Some talents work particularly well together:

\subsection*{Casting Synergies}

\begin{itemize}
ℹtem Spellcraft + Spell Shaping: Flexible, precise freeform casting
ℹtem Elemental Mastery + Arcane Dominance: Powerful, controlled elemental effects
ℹtem Ritual Mastery + Spellcraft: Reduced risk on both freeform and ritual casting
\end{itemize}

\subsection*{Social Synergies}

\begin{itemize}
ℹtem Silver Tongue + Command Presence: Charm or command with equal force
ℹtem Read Emotions + Network Builder: Understand and leverage social connections
ℹtem Familiar + Social Talents: Patron−enhanced social abilities
\end{itemize}

\subsection*{Exploration Synergies}

\begin{itemize}
ℹtem Keen Senses + Trackless Step: Find others while leaving no trace
ℹtem Wilderness Lore + Urban Navigation: Comfortable in all environments
ℹtem Familiar + Exploration Talents: Patron−guided exploration
\end{itemize}

\section*{Talent Limitations and Balance} ∈dex{Talents!limitations}

\subsection*{Usage Restrictions}

Most talents have limits to maintain game balance:
\begin{itemize}
ℹtem Per scene: Common for strong actives
ℹtem Per session: Reserved for swingy effects
ℹtem Resource cost: Some require spending Boons or generating Obligation
ℹtem Position requirements: May require specific narrative circumstances
\end{itemize}

\subsection*{Prerequisite Systems}

Advanced talents require meeting certain conditions:
\begin{itemize}
ℹtem Attribute minimums: e.g., Spirit 4, Wits 3
ℹtem Skill requirements: Specific skills at set levels
ℹtem Previous talents: Foundational picks first (Familiar required for Patron features)
ℹtem Tier requirements: Character advancement level
\end{itemize}

\section*{Building Your Talent Set} ∈dex{Talents!building}

\subsection*{Early Game (0−−40 XP)}

Focus on essential access and basic capabilities:
\begin{itemize}
ℹtem 1−−2 access talents (Spellcraft, Familiar)
ℹtem 2−−3 basic talents for reliability
ℹtem Save XP for major access requirements
ℹtem Choose talents that work with your core concept
\end{itemize}

\subsection*{Mid Game (41−−90 XP)}

Develop your specialization:
\begin{itemize}
ℹtem Major access talents (Codex, Patron's Symbol)
ℹtem 2−−3 synergistic power talents
ℹtem Balance active and passive picks
ℹtem Plan for prestige abilities
\end{itemize}

\subsection*{Late Game (91+ XP)}

Achieve mastery:
\begin{itemize}
ℹtem 1−−2 prestige talents defining your apex
ℹtem Picks that create legacy effects
ℹtem Talents that benefit the whole party
ℹtem Prepare for campaign−defining challenges
\end{itemize}

\section*{Talent Customization} ∈dex{Talents!customization}

Work with your Game Master to create custom talents:
\begin{itemize}
ℹtem Based on story events: Reflect character experiences
ℹtem Balanced costs: Match similar scope to existing talents (2/4/6+ XP)
ℹtem Clear prerequisites: Define requirements clearly
ℹtem Mechanical clarity: Define activation, effects, and limits
\end{itemize}

\section*{Talents and Group Dynamics} ∈dex{Talents!group dynamics}

Consider how your talents complement the party:
\begin{itemize}
ℹtem Fill gaps: Cover party weaknesses in magical capabilities
ℹtem Synergize: Coordinate with other players' magical approaches
ℹtem Avoid overlap: Don't duplicate another character's access path
ℹtem Support role: Talents that help the whole group manage magical risks
\end{itemize}

\section*{Talent Respecification} ∈dex{Talents!respecification}

If your character concept changes, you may respec talents:
\begin{itemize}
ℹtem GM approval required: Discuss proposed changes
ℹtem Downtime cost: Represent retraining (typically 1 downtime period)
ℹtem Story justification: Explain the change in−narrative
ℹtem Limited frequency: Typically once per major story arc
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Talent Selection Guide, fonttitle=\bfseries]
\textbf{Early Game (0−−40 XP):}
\begin{itemize}
ℹtem 1−−2 access talents (2 XP each)
ℹtem 2−−4 basic talents (2 XP each)
ℹtem Focus on essential capabilities
\end{itemize}

\textbf{Mid Game (41−−90 XP):}
\begin{itemize}
ℹtem 1−−2 major talents (4 XP each)
ℹtem 1−−2 advanced talents (4−−6 XP each)
ℹtem Plan for prestige prerequisites
\end{itemize}

\textbf{Late Game (91+ XP):}
\begin{itemize}
ℹtem 1−−2 prestige talents (6+ XP each)
ℹtem Campaign−defining capabilities
ℹtem Party−supporting abilities
\end{itemize}

\textbf{Remember:} Talents should reflect your character's story and magical growth.
\end{tcolorbox}

\section*{Practical Talent Examples} ∈dex{Talents!practical examples}

\subsection*{Example 1: The Caster}

\begin{itemize}
ℹtem Spellcraft (6 XP) −−− Essential access to freeform casting
ℹtem Spell Shaping (4 XP) −−− Modify spell parameters
ℹtem Elemental Mastery (6 XP) −−− Reduce casting risks
ℹtem Arcane Dominance (6 XP) −−− Overpower opposing magic
ℹtem Total: 18 XP invested in casting capabilities
\end{itemize}

\subsection*{Example 2: The Runekeeper}

\begin{itemize}
ℹtem Familiar (2 XP) −−− Access to Patron features
ℹtem Codex (4 XP) −−− Full Rites access
ℹtem Ritual Mastery (4 XP) −−− Reduced ritual risks
ℹtem Obligation Master (8 XP) −−− Better debt management
ℹtem Total: 18 XP invested in Pact magic
\end{itemize}

\subsection*{Example 3: The Invoker}

\begin{itemize}
ℹtem Patron's Symbol (4 XP) −−− Ritual access to Patron
ℹtem Efficient Invocation (4 XP) −−− Faster rituals
ℹtem Crack Specialist (6 XP) −−− Reduced instant cast costs
ℹtem Ritual Mastery (4 XP) −−− Reduced backlash
ℹtem Total: 18 XP invested in ritual magic
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Final Note, fonttitle=\bfseries]
The best talents are those that fit your magical concept and table playstyle. Choose abilities you'll enjoy using, that create interesting consequences, and that contribute to your character's unfolding story through the lens of risk and consequence that defines Fate's Edge magic.
\end{tcolorbox}

\section*{Narrative−Heavy Talent Options} ∈dex{Narrative−Heavy Gameplay!Talents}

For groups that prefer strong narrative focus in talent use, consider these optional approaches:

\textbf{Story−Driven Talents:} Instead of mechanical bonuses, some talents can provide narrative permissions or story effects. "Courtly Grace" might allow you to navigate noble society without rolls, while "Wild Empathy" lets you communicate with animals through roleplay rather than dice.

\textbf{Collaborative Talent Activation:} Players can describe how their talents work in the fiction, with GM approval, rather than relying solely on mechanical triggers. A "Master Strategist" might narrate how they reposition allies through clever tactics rather than just declaring the mechanical effect.

\textbf{Talent as Character Development:} Use talent acquisition as opportunities for character growth and backstory development, allowing players to narrate how their characters learned new abilities through significant story moments.

\textbf{Flexible Talent Interpretation:} Focus on the thematic effects of talents rather than strict mechanical applications. A "Weapon Mastery" talent might manifest differently depending on the weapon and situation, with the GM and player collaborating on the specific benefits.

\chapter{Assets and Followers} łabel{ch:assets}

Your character's influence extends beyond personal capabilities through Assets and Followers. These represent worldly possessions, connections, and allies that can solve problems, provide assistance, and shape the narrative.

\section*{Understanding Assets and Followers} ∈dex{Assets}∈dex{Followers}

\subsection*{Key Differences}

\begin{description}
ℹtem[Assets:] Off−screen resources that solve problems between scenes.
ℹtem[Followers:] On−screen allies who assist during gameplay.
ℹtem[Assets change:] The fictional situation before you arrive.
ℹtem[Followers act:] Alongside you in the moment.
\end{description}

\subsection*{Management Requirements}

Both require maintenance and carry risks:
\begin{itemize}
ℹtem Regular upkeep costs (XP or downtime).
ℹtem Vulnerability to complications and attacks.
ℹtem Narrative consequences for misuse or neglect.
\end{itemize}

\section*{Assets System} ∈dex{Assets!system}

Assets are possessions, properties, or resources you control.

\subsection*{Asset Types and Costs}

\begin{center}
\begin{tabular}{|l|r|l|}
ℎline
\textbf{Type} & \textbf{XP Cost} & \textbf{Establishment Time} \\
ℎline
Minor & 4 XP & 1 day \\
Standard & 8 XP & 1 week \\
Major & 12 XP & 1 month \\
ℎline
\end{tabular}
\end{center}

\subsection*{Asset Examples}

\begin{description}
ℹtem[Minor Assets] Small shop, safehouse, minor title, basic workshop.
ℹtem[Standard Assets] Noble title, guild membership, trading post, spy network.
ℹtem[Major Assets] Fortress, city license, major enterprise, regional influence.
\end{description}

\subsection*{Using Assets}

Assets provide benefits in different ways:

⫕section*{Free Off−Screen Use}

Each asset has a specific off−screen effect you can use once per session:
\begin{itemize}
ℹtem Safehouse: Provide secure lodging for the party.
ℹtem Spy Network: Gather basic intelligence about a location.
ℹtem Workshop: Repair or create simple items between adventures.
ℹtem Trading Post: Acquire common goods at better prices.
\end{itemize}

⫕section*{Boon Activation}

Spend 1 Boon to use an asset dramatically during a scene:
\begin{itemize}
ℹtem Safehouse: Suddenly reveal a hidden escape route.
ℹtem Spy Network: Produce crucial information at a critical moment.
ℹtem Workshop: Create an improvised solution to an immediate problem.
ℹtem Trading Post: Call in a favor from a business contact.
\end{itemize}

⫕section*{XP Activation}

Spend 2 XP to use an asset's off−screen effect outside your normal allowance:
\begin{itemize}
ℹtem Emergency use when you've already used your free activation.
ℹtem Additional uses during downtime periods.
ℹtem Special circumstances requiring extra asset support.
\end{itemize}

\section*{Asset Conditions} ∈dex{Assets!conditions}

Assets have condition states affecting their usefulness:

\subsection*{Condition Levels}

\begin{description}
ℹtem[Maintained] Fully functional, no penalties.
ℹtem[Neglected] −1 die when used; requires attention.
ℹtem[Compromised] Unavailable until repaired or recovered.
\end{description}

\subsection*{Maintenance Requirements}

\begin{itemize}
ℹtem Regular Upkeep: Two options per SRD §21.2:
\begin{itemize}
ℹtem Efficient (Higher XP, Less Time): Pay Upkeep XP = max(1, XP Acquisition)/3, minimal effort
ℹtem Intensive (Lower XP, More Time): Pay 1 XP, dedicated downtime action
\end{itemize}
ℹtem Neglect: Assets deteriorate if not maintained.
ℹtem Recovery: Compromised assets require significant effort to restore.
\end{itemize}

\section*{Followers System} ∈dex{Followers!system}

Followers are characters who assist you directly.

\subsection*{Follower Capability Ratings}

Followers are rated by Capability (Cap) from 1 to 5:

\begin{center}
\begin{tabular}{|c|l|}
ℎline
\textbf{Cap} & \textbf{Description} \\
ℎline
1 & Novice helper, basic assistance \\
2 & Competent assistant, reliable support \\
3 & Skilled specialist, valuable aid \\
4 & Expert ally, significant capability \\
5 & Master companion, exceptional ability \\
ℎline
\end{tabular}
\end{center}

\subsection*{Follower Costs}

\begin{itemize}
ℹtem XP Cost: Capability squared (Cap$²$).
ℹtem Example: Cap 3 follower costs 3$²$ = 9 XP.
ℹtem Recruitment: 1−−3 days downtime to find and brief.
ℹtem Limits: The GM may set maximum followers based on story.
\end{itemize}

\subsection*{Follower Types}

\begin{description}
ℹtem[Combat Allies] Warriors, guards, mercenaries.
ℹtem[Technical Experts] Craftspeople, engineers, specialists.
ℹtem[Social Contacts] Informants, diplomats, agents.
ℹtem[Specialists] Unique capabilities like magic or stealth.
\end{description}

\section*{Using Followers} ∈dex{Followers!usage}

\subsection*{Assistance in Scenes}

Followers can help with your actions:
\begin{itemize}
ℹtem Assist Dice: Add dice equal to min(Cap, relevant skill).
ℹtem Maximum Bonus: +3 dice total from all sources.
ℹtem Cost: Spend 1 Boon or 1 Stress to add +1 die (max +3 from assists).
ℹtem One Helper: Only one follower can assist per action.
\end{itemize}

\subsection*{Independent Actions}

Once per scene (party−wide), a follower can take a small action:
\begin{description}
ℹtem[Scout & Signal] Change an ally's next action to Controlled position.
ℹtem[Distract & Draw] Reduce a threat timer by 1 segment.
ℹtem[Fetch & Carry] Move an object through danger safely.
\end{description}

\subsection*{Cost of Independent Actions}

\begin{itemize}
ℹtem Mark +1 Exposure (attention or stress), or
ℹtem Take Harm 1 (injury or trauma).
ℹtem Cannot be used if the follower is already Compromised.
\end{itemize}

\section*{Follower Conditions} ∈dex{Followers!conditions}

Followers track two condition types:

\subsection*{Exposure}

Represents attention, stress, or narrative pressure:
\begin{itemize}
ℹtem Gains: From independent actions, dangerous situations, complications.
ℹtem Effects: Increased risk, reduced effectiveness, attention from enemies.
ℹtem Recovery: Downtime activities, careful management.
\end{itemize}

\subsection*{Harm}

Represents injury, trauma, or damage:
\begin{itemize}
ℹtem Gains: From combat, accidents, enemy attacks.
ℹtem Effects: Penalties to assistance, possible incapacity.
ℹtem Recovery: Medical care, rest, magical healing.
\end{itemize}

\subsection*{Condition States}

\begin{description}
ℹtem[Maintained] Ready and reliable, full capability.
ℹtem[Neglected] Needs attention, −1 die to assistance.
ℹtem[Compromised] Unavailable: captured, defected, lost, or incapacitated.
\end{description}

\section*{Follower Risks} ∈dex{Followers!risks}

Using followers carries significant risks:

\subsection*{Complication Targeting}

When the GM spends 2+ Story Beats on an action where you have assistance:
\begin{itemize}
ℹtem The follower may face consequences instead of you.
ℹtem Could be injury, capture, betrayal, or other complications.
ℹtem Fictionally appropriate to the situation.
\end{itemize}

\subsection*{Off−Screen Capability}

Once per downtime, a Cap 5 follower can solve a significant problem:
\begin{itemize}
ℹtem But generates 1 Story Beat for the party.
ℹtem The GM describes how this creates story consequences.
ℹtem Useful for emergencies but costly.
\end{itemize}

\section*{Upkeep and Maintenance} ∈dex{Assets!upkeep}∈dex{Followers!upkeep}

Both assets and followers require regular maintenance.

\subsection*{Asset Upkeep}

Two options per SRD §21.2:
\begin{itemize}
ℹtem Option 1 − Efficient (Higher XP, Less Time):
\begin{itemize}
ℹtem Cost: Pay Upkeep XP = max(1, AcquisitionXP )/3
ℹtem Time: Minimal effort
\end{itemize}
ℹtem Option 2 − Intensive (Lower XP, More Time):
\begin{itemize}
ℹtem Cost: Pay 1 XP
ℹtem Time: Dedicated downtime action with significant personal involvement
\end{itemize}
ℹtem Failure to Pay: Asset becomes Neglected (or Compromised if already Neglected)
\end{itemize}

\subsection*{Follower Upkeep}

Two options per SRD §21.2:
\begin{itemize}
ℹtem Option 1 − Efficient:
\begin{itemize}
ℹtem Cost: Pay Upkeep XP = max(1, Cap$²$)/3
ℹtem Time: Minimal effort
\end{itemize}
ℹtem Option 2 − Intensive:
\begin{itemize}
ℹtem Cost: Pay 1 XP
ℹtem Time: Dedicated downtime action with significant personal involvement
\end{itemize}
ℹtem Failure to Pay: Follower becomes Wary (or Seized if already Wary)
\end{itemize}

\section*{Strategic Considerations} ∈dex{Assets!strategy}∈dex{Followers!strategy}

\subsection*{When to Invest in Assets}

\begin{itemize}
ℹtem You need reliable off−screen capabilities.
ℹtem Your character concept involves wealth or influence.
ℹtem The party lacks certain logistical support.
ℹtem You want to build long−term influence.
\end{itemize}

\subsection*{When to Invest in Followers}

\begin{itemize}
ℹtem You need on−screen assistance.
ℹtem Your character works better with support.
ℹtem The party needs specific capabilities you lack.
ℹtem You want character−driven story opportunities.
\end{itemize}

\subsection*{Balance Recommendations}

\begin{center}
\begin{tabular}{|l|l|}
ℎline
\textbf{Path} & \textbf{Investment} \\
ℎline
Personal Path & 0−−10% assets/followers \\
Balanced Path & 15−−░ assets/followers \\
Influencer Path & 35−−55% assets/followers \\
ℎline
\end{tabular}
\end{center}

\section*{Loyalty and Relationships} ∈dex{Followers!loyalty}

\subsection*{Loyalty Levels}

Optional system for tracking follower loyalty:
\begin{description}
ℹtem[Wary] Cautious, may leave if pressured; +1 XP upkeep cost.
ℹtem[Steady] Reliable, standard performance; normal upkeep.
ℹtem[Devoted] Loyal, may sacrifice; can convert one major complication to a minor setback per arc.
\end{description}

\subsection*{Building Loyalty}

\begin{itemize}
ℹtem Fair treatment and respect.
ℹtem Sharing rewards and successes.
ℹtem Protecting followers from harm.
ℹtem Honoring agreements and promises.
\end{itemize}

\subsection*{Losing Loyalty}

\begin{itemize}
ℹtem Mistreatment or disrespect.
ℹtem Unreasonable demands or risks.
ℹtem Broken promises or betrayal.
ℹtem Consistent neglect.
\end{itemize}

\section*{Advanced Follower Management} ∈dex{Followers!advanced}

\subsection*{Follower Groups}

For multiple similar followers, you can manage them as a group:
\begin{itemize}
ℹtem Single Rating: Treat as one entity with combined capability.
ℹtem Condition Tracking: Group shares exposure and harm.
ℹtem Maintenance: Single upkeep cost for the group.
ℹtem Risks: Problems affect the entire group.
\end{itemize}

\subsection*{Follower Advancement}

Followers can improve over time:
\begin{itemize}
ℹtem Experience: Gain capability through successful assistance.
ℹtem Training: Spend XP to improve follower capabilities.
ℹtem Equipment: Better gear can enhance effectiveness.
ℹtem Limits: Followers typically cap at lower levels than PCs.
\end{itemize}

\section*{Risk Management} ∈dex{Assets!risk management}∈dex{Followers!risk management}

\subsection*{Asset Risks}

\begin{itemize}
ℹtem Financial: Assets can be costly to maintain.
ℹtem Security: Assets can be attacked or stolen.
ℹtem Attention: Valuable assets draw unwanted notice.
ℹtem Dependency: Over−reliance can be problematic.
\end{itemize}

\subsection*{Follower Risks}

\begin{itemize}
ℹtem Safety: Followers can be harmed or captured.
ℹtem Loyalty: Followers may betray or leave.
ℹtem Attention: Followers can draw enemy interest.
ℹtem Morale: Followers have needs and limits.
\end{itemize}

\subsection*{Mitigation Strategies}

\begin{itemize}
ℹtem Diversification: Don't put all resources in one place.
ℹtem Security: Protect valuable assets and followers.
ℹtem Relationships: Maintain good terms with your people.
ℹtem Contingencies: Have backup plans for losses.
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Assets and Followers Quick Reference, fonttitle=\bfseries]
\textbf{Assets:}
\begin{itemize}
ℹtem Minor: 4 XP | Standard: 8 XP | Major: 12 XP
ℹtem Free off−screen use: once per session
ℹtem Boon activation: spend 1 Boon for scene impact
ℹtem Conditions: Maintained → Neglected → Compromised
\end{itemize}

\textbf{Followers:}
\begin{itemize}
ℹtem Cost: Cap$²$ XP
ℹtem Assistance: + min(Cap, skill) dice (max +3 from all sources)
ℹtem Independent action: once per scene (party−wide)
ℹtem Conditions: Exposure and Harm tracks
\end{itemize}

\textbf{Upkeep Options:}
\begin{itemize}
ℹtem Efficient: max(1, Cost)/3 XP, minimal time
ℹtem Intensive: 1 XP, dedicated downtime action
\end{itemize}
\end{tcolorbox}

\section*{Practical Examples} ∈dex{Assets!examples}∈dex{Followers!examples}

\subsection*{Asset Example: The Safehouse}

\begin{itemize}
ℹtem Type: Minor Asset (4 XP)
ℹtem Free Use: Secure lodging, basic supplies between adventures.
ℹtem Boon Activation: Reveal a hidden escape route during pursuit.
ℹtem Upkeep: Option 1: 2 XP (4/3 rounded up) or Option 2: 1 XP + downtime action.
ℹtem Risks: Discovery by enemies, maintenance costs.
\end{itemize}

\subsection*{Follower Example: The Scout}

\begin{itemize}
ℹtem Capability: 3 (9 XP cost)
ℹtem Assistance: +3 dice on tracking and survival rolls.
ℹtem Independent Action: Scout ahead to improve party position.
ℹtem Upkeep: Option 1: 3 XP (9/3) or Option 2: 1 XP + downtime action.
ℹtem Risks: Injury in dangerous scouting; disloyalty if mistreated.
\end{itemize}

\subsection*{Combination Example: The Merchant}

\begin{itemize}
ℹtem Assets: Trading post (8 XP), caravan (4 XP) −−− 12 XP total
ℹtem Followers: Cap 2 guards (4 XP each = 8 XP), Cap 3 factor (9 XP) −−− 17 XP total
ℹtem Total Investment: 29 XP in assets and followers
ℹtem Upkeep (Efficient Option): Assets 4 XP + Followers 6 XP = 10 XP per downtime period
ℹtem Benefits: Trade income, transport, protection, business contacts
ℹtem Risks: Competition, bandit attacks, employee issues, regulatory attention
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Remember, fonttitle=\bfseries]
Assets and followers can greatly expand your capabilities, but they require careful management and carry significant risks. Invest wisely based on your character concept and the needs of your group. The SRD provides flexible upkeep options to suit different play styles and campaign pacing.
\end{tcolorbox}

\section*{Narrative−Heavy Asset and Follower Options} ∈dex{Narrative−Heavy Gameplay!Assets and Followers}

For groups that prefer strong narrative focus in asset and follower management, consider these optional approaches:

\textbf{Story−Driven Upkeep:} Instead of tracking XP costs for upkeep, the GM can introduce narrative complications that require attention. A neglected asset might attract unwanted attention, while a neglected follower might request a favor or special treatment.

\textbf{Collaborative Management:} Players can describe how they maintain their assets and followers through roleplay rather than mechanical upkeep costs. A well−described scene of tending to a workshop or bonding with followers can fulfill maintenance requirements.

\textbf{Asset and Follower as Character Development:} Use asset and follower management as opportunities for character growth and backstory development, allowing players to narrate how their relationships and holdings evolve through significant story moments.

\textbf{Flexible Condition Tracking:} Focus on the narrative implications of asset and follower conditions rather than strict mechanical penalties. A "Neglected" asset might still function but with interesting complications, while a "Compromised" asset might require creative solutions rather than just XP investment.

\chapter{World Interaction} łabel{ch:world}

In Fate's Edge, the world is not a backdrop−−−it's a partner in the conversation. Dikes groan under black rain in Viterra, clan horns answer across Acasia's ridgelines, Ecktoria's marble halls echo with careful words, and Kahfagia's pilots read storms by taste. Wherever you go, place, culture, and pressure push back.

\section*{Game Structure and Time} ∈dex{Game Structure}

Understanding how time works in Fate's Edge helps you navigate both the mechanical and narrative flow of play.

\subsection*{Basic Units}

\begin{description}
ℹtem[Scene] The basic unit of narrative play, covering a specific situation or conflict (Some Time to Significant Time). Resolves a particular question or challenge.
ℹtem[Player Turn (Beat)] An individual player's action within a scene: Declare action → GM sets position → roll → resolve outcome → manage consequences.
ℹtem[Round] Simultaneous or near−simultaneous actions within a scene (primarily for combat), representing a few seconds of real time.
ℹtem[Session] One complete game session (typically 3−−6 hours), containing 2−−4 major scenes and resolving significant narrative progress.
ℹtem[Downtime] The narrative time between scenes, used for recovery, advancement, and off−screen activities. Measured in days, weeks, or months depending on fiction.
ℹtem[Campaign] Entire story arc (6−−20+ sessions) with major character development and lasting consequences.
\end{description}

\section*{Movement and Positioning} ∈dex{Movement}

Space is tracked with range bands and Position.

\subsection*{Range Bands}

\begin{description}
ℹtem[Close] Touching distance: grapples, knife−work, hand on a relic.
ℹtem[Near] Same room/yard/deck; a rush away.
ℹtem[Far] Same site but distant; requires route or time to reach.
ℹtem[Absent] Off−screen; requires scene change or significant effort to interact.
\end{description}

\subsection*{Movement Actions}

\begin{itemize}
ℹtem Move: Shift one range band as a beat.
ℹtem Dash: Shift two bands as your full action (terrain may require a roll).
ℹtem Melee Flag: Mark when two parties are in Near range and directly engaged in combat.
\end{itemize}

\subsection*{Position States}

\begin{description}
ℹtem[Controlled] You have cover, leverage, or ritual footing. Failure still leaves options.
ℹtem[Controlled] Standard case: exposed lanes, rivals near, watchful eyes. Failure has teeth, but not ruin.
ℹtem[Desperate] Bad ground, bad odds, bad timing. Failure is severe; success may bring extra XP.
\end{description}

\subsection*{Position Shifting}

\begin{itemize}
ℹtem GM can spend 1 SB to worsen Position by one step.
ℹtem Player can spend 1 Boon to improve Position by one step (once per action).
ℹtem Narrative triggers (flanking, reinforcements, etc.) can shift Position without cost.
\end{itemize}

\section*{Travel Framework} ∈dex{Travel}

Travel abstracts distance into legs with tension and color rather than miles and meal counts. Each leg has a Travel Timer and draws on a regional deck to seed fiction.

\subsection*{Travel Process}

\begin{enumerate}
ℹtem Set the Leg: Name origin and destination; start a Travel Timer (4−10 segments based on difficulty).
ℹtem Draw Prompts: Draw up to one card from each suit to establish terrain, people, pressures, and leverage.
ℹtem Assign Roles: Players take on travel roles (Guide, Scout, Quartermaster, Watch) to contribute actions.
ℹtem Play the Leg: Players take actions to advance the timer or mitigate complications. GM spends SB from rolls showing 1s to introduce hazards.
ℹtem Resolve: When the timer fills, you arrive−−−changed by the journey.
\end{enumerate}

\subsection*{Using Assets and Followers During Travel}

\begin{itemize}
ℹtem Assets: Spend 1 Boon to activate an asset for dramatic effect during travel (reveal hidden path, call for emergency aid, etc.).
ℹtem Followers: Assign followers to travel roles for bonuses. A Cap 3 Scout follower adds +3 to navigation rolls, for example.
ℹtem Independent Actions: Once per travel leg, a follower can take an independent action (scout ahead, secure supplies, etc.) at the cost of Exposure or Harm.
ℹtem Off−Screen Solutions: High−Cap followers (4−5) can solve significant travel problems once per downtime, but generate 1 SB for the party.
\end{itemize}

\subsection*{Regional Travel Decks}

Each major region has a themed prompt list or card table:
\begin{description}
ℹtem[Viterra] Fen causeways, dike−brotherhoods, crown law.
ℹtem[Acasia] Border−lace titles, ruined towers, clan tempers.
ℹtem[Ecktoria] Imperial roads, precinct gates, temple schedules.
ℹtem[Ubral] Stone passes, toll−cloisters, ghosted fields.
ℹtem[Kahfagia] Current maps, pilot−mirrors, storm lanes.
ℹtem[Aelinnel] Mist paths, bell−mounds, spirit ways.
\end{description}

\subsection*{Travel Complications}

\begin{itemize}
ℹtem Hazards: Weather, terrain challenges, wildlife encounters.
ℹtem Social: Border checks, local politics, cultural misunderstandings.
ℹtem Supplies: Food shortages, equipment failure, resource management.
ℹtem Pursuit: Being followed, hunted, or racing against time.
\end{itemize}

\section*{Narrative Time} ∈dex{Time}

Time is measured by importance rather than duration.

\begin{description}
ℹtem[A Moment] A glance, a strike, a whisper over a law−stone.
ℹtem[Some Time] A skirmish, a negotiation, a careful climb.
ℹtem[Significant Time] Hours of march, rites, audits, stakeouts.
ℹtem[Days] Drills, recoveries, research, roadwork.
\end{description}

\section*{Social Interactions} ∈dex{Social Interactions}

Social scenes use the same engine with cultural color.

\subsection*{Cultural Skill Emphases}

\begin{description}
ℹtem[Viterra] Rapport with parishes; Sway for markets; Command under writ.
ℹtem[Acasia] Rapport for kin−bridges; Command with banner−rights; Deceive risks honor timers.
ℹtem[Ecktoria] Sway in salons; Deceive at court; Perform in temple fora.
ℹtem[Kahfagia] Rapport aboard; Sway at piers; Command on a storming deck.
\end{description}

\subsection*{Social Stakes & Timers}

\begin{itemize}
ℹtem Alliance Timer (Viterra): Parishes and guilds come to your side.
ℹtem Honor Timer (Acasia): Feasts, oaths, wyrd−−−trust builds (or frays).
ℹtem Bureau Timer (Ecktoria): Stamps, seals, approvals−−−delay is pressure.
ℹtem Trust Timer (Kahfagia): Pilots and crews extend favors and routes.
\end{itemize}

\section*{Supply and Resources} ∈dex{Resources}

Track scarcity with a Supply Timer shared by the party's expedition.

\subsection*{Segments}

\begin{center}
\begin{tabular}{|c|l|l|}
ℎline
\textbf{Segments} & \textbf{State & Effects} & \textbf{Notes} \\
ℎline
0 (Full) & Well−provisioned; no penalty. & \\
2 (Low) & Minor frictions; −1 to resource checks. & \\
3 (Dangerous) & Each PC gains Fatigue 1. & \\
4 (Empty) & Severe penalties; desperate measures. & \\
ℎline
\end{tabular}
\end{center}

\section*{Engaging the World−−−Player Actions} ∈dex{Player Actions}

\begin{itemize}
ℹtem Scout & Signal: A follower can make the next travel action Controlled (mark Exposure or Harm 1 on them).
ℹtem Local Color: Briefly state what locals notice about you; GM offers a small fictional edge or a tempting timer−−−choose.
ℹtem Mark the Map: On arrival, declare one change to the fiction (new ford, patron's shrine, toll−skip). GM may attach a minor timer as cost.
ℹtem Asset Activation: Spend 1 Boon to activate an asset dramatically during a scene.
ℹtem Follower Assistance: Have a follower assist your actions for bonus dice (max +3 from all sources).
\end{itemize}

\section*{Summary} ∈dex{World Interaction!summary}

The world has opinions. Movement is timers and color, position rises and sinks with weather and words, and every suit you draw speaks in a regional accent. Ask the land for a favor−−−then pay it back on the road.

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Remember, fonttitle=\bfseries]
Every interaction with the world is an opportunity. Use your assets, deploy your followers, and engage with the setting actively. The world responds to your choices, and every journey changes both you and the places you pass through.
\end{tcolorbox}

\chapter{Example Character Concepts} łabel{ch:examples}

This chapter presents example character concepts to illustrate how the game's systems can create diverse and interesting heroes. These are examples only−−−not prescriptive templates or exhaustive lists. Use them for inspiration, as pre−generated characters, or as starting points for your own unique creations.

\section*{Important Disclaimer} ∈dex{Character Creation!disclaimer}

These examples are provided for illustrative purposes only. They demonstrate how the game's mechanics can support different character archetypes and play styles. You are encouraged to:
\begin{itemize}
ℹtem Modify these concepts to fit your preferences
ℹtem Create completely original characters
ℹtem Mix and match elements from different examples
ℹtem Work with your Game Master to develop unique concepts
\end{itemize}

The game system is designed to support a wide variety of character types beyond these examples.

\section*{How to Use These Examples} ∈dex{Character Creation!usage}

Each concept includes:
\begin{itemize}
ℹtem Concept Overview: Narrative identity and role
ℹtem Mechanical Foundation: Suggested starting capabilities
ℹtem Play Style: How the character typically engages with challenges
ℹtem Development Path: Potential growth directions
ℹtem Story Hooks: Plot opportunities for the Game Master
ℹtem Build Blocks: A 32 XP starting build, plus an optional 36XP variant using Bonds/Complications (+4 XP)
\end{itemize}

\section*{1. The Guardian} ∈dex{Character Examples!Guardian}

\textbf{Concept:} A protector who stands between danger and those they've sworn to defend. Steel in hand, vow in heart.

\textbf{Typical Inspiration:} Paladins, knights, bodyguards, sworn shields

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Body, Spirit
ℹtem Skills: Melee, Athletics, Command
ℹtem Talents: Defensive stance, protective instincts
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Frontline combat and protection
ℹtem Drawing attention away from allies
ℹtem Using presence and authority to control situations
ℹtem Taking risks to protect others
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Increase defensive capabilities
ℹtem Develop leadership skills
ℹtem Acquire better protective gear
ℹtem Learn area control abilities
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem Who or what are they protecting?
ℹtem What oath or duty drives them?
ℹtem What happens if they fail in their protection?
ℹtem What personal costs do they bear for their role?
\end{itemize}

\textbf{Build Blocks. Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes (Cost = rating × 3 XP): Body 3 (9), Spirit 2 (6), Wits 1 (3), Presence 1 (3) → 21 XP
ℹtem Skills (Cost = level × 2 XP): Melee 2 (4), Athletics 1 (2), Command 1 (2) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Talent: Combat Reflexes (5 XP) using banked 1 + 4 = 5 XP
ℹtem Revised Total: 36XP
\end{itemize}

\section*{2. The Scholar} ∈dex{Character Examples!Scholar}

\textbf{Concept:} A seeker of knowledge who uses information as power. Candlesmoke, marginalia, and dangerous truths.

\textbf{Typical Inspiration:} Wizards, sages, researchers, historians

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Wits, Spirit
ℹtem Skills: Lore, Investigation, Arcana
ℹtem Talents: Quick Study, Research Mastery
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Information gathering and analysis
ℹtem Solving puzzles and mysteries
ℹtem Using knowledge to gain advantages
ℹtem Researching solutions between adventures
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Specialize in specific knowledge areas
ℹtem Develop magical or technical capabilities
ℹtem Build research networks
ℹtem Create unique inventions or discoveries
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem What knowledge are they seeking?
ℹtem What dangerous information might they uncover?
ℹtem How do they handle forbidden knowledge?
ℹtem Who opposes their research?
\end{itemize}

\textbf{Build Blocks. Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes: Wits 3 (9), Spirit 2 (6), Body 1 (3), Presence 1 (3) → 21 XP
ℹtem Skills: Lore 2 (4), Investigation 1 (2), Arcana 1 (2) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Talent: Research Mastery (5 XP) using banked 1 + 4 = 5 XP
ℹtem Revised Total: 36XP
\end{itemize}

\section*{3. The Scout} ∈dex{Character Examples!Scout}

\textbf{Concept:} A wilderness expert who navigates dangerous territories. Quiet footfalls, hawk eyes, and the long road.

\textbf{Typical Inspiration:} Rangers, hunters, trackers, explorers

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Wits, Body
ℹtem Skills: Survival, Stealth, Perception
ℹtem Talents: Wilderness Lore, Keen Senses
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Scouting ahead and gathering intelligence
ℹtem Wilderness survival and navigation
ℹtem Ambush and skirmish tactics
ℹtem Finding paths and resources
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Improve stealth and tracking abilities
ℹtem Develop animal companions or allies
ℹtem Master specific environments
ℹtem Learn advanced survival techniques
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem What uncharted territory are they exploring?
ℹtem What secrets have they discovered in the wild?
ℹtem How do they balance civilization and wilderness?
ℹtem What threats have they encountered beyond settled lands?
\end{itemize}

\textbf{Build Blocks. Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes: Wits 3 (9), Body 2 (6), Spirit 1 (3), Presence 1 (3) → 21 XP
ℹtem Skills: Survival 2 (4), Stealth 2 (4) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Asset: Hidden Cache (Minor Asset, 4 XP) using banked 1 + 4 = 5 XP
ℹtem Revised Total: 36XP
\end{itemize}

\section*{4. The Diplomat} ∈dex{Character Examples!Diplomat}

\textbf{Concept:} A negotiator who resolves conflicts through words and influence. A smile for the foyer, steel for the parlor.

\textbf{Typical Inspiration:} Bards, ambassadors, merchants, politicians

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Presence, Wits
ℹtem Skills: Sway, Investigation, Lore
ℹtem Talents: Silver Tongue, Read Emotions
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Social interaction and negotiation
ℹtem Gathering information through contacts
ℹtem Resolving conflicts without violence
ℹtem Building alliances and relationships
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Expand social influence and networks
ℹtem Develop economic or political power
ℹtem Learn cultural specialties
ℹtem Master manipulation or inspiration techniques
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem What major conflict are they trying to resolve?
ℹtem What alliances have they built or broken?
ℹtem How do they handle betrayal or failed negotiations?
ℹtem What personal relationships affect their diplomacy?
\end{itemize}

\textbf{Build Blocks. Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes: Presence 3 (9), Wits 2 (6), Spirit 1 (3), Body 1 (3) → 21 XP
ℹtem Skills: Sway 2 (4), Investigation 1 (2), Lore 1 (2) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Talent: Silver Tongue (3 XP) and Skill: Lore +1 (now 2) for 2 XP using banked 1 + 4 = 5 XP
ℹtem Revised Total: 36XP
\end{itemize}

\section*{5. The Specialist} ∈dex{Character Examples!Specialist}

\textbf{Concept:} An expert with unique capabilities beyond typical roles. The right tool, the right touch, at the right time.

\textbf{Typical Inspiration:} Artisans, healers, engineers, spies

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Varies by specialty (often Wits or Body)
ℹtem Skills: One specialty at focus, plus two support skills
ℹtem Talents: Unique techniques that unlock niche actions
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Solving problems with unique expertise
ℹtem Creating or repairing specialized items
ℹtem Providing services others cannot
ℹtem Using niche knowledge for advantage
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Master their specialty area
ℹtem Develop related capabilities
ℹtem Build reputation and clientele
ℹtem Create unique inventions or methods
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem What makes their specialty unique or valuable?
ℹtem How did they acquire their special skills?
ℹtem What problems require their specific expertise?
ℹtem Who seeks to control or exploit their abilities?
\end{itemize}

\textbf{Build Blocks (Artificer example). Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes: Wits 3 (9), Body 2 (6), Presence 1 (3), Spirit 1 (3) → 21 XP
ℹtem Skills: Craft 2 (4), Mechanics 2 (4) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Talent: Technical Expert (6 XP) − need 6 XP but have 5 XP available (1 banked + 4 from Bonds/Complications)
ℹtem Alternative: Add Talent: Quick Study (3 XP) and bank 2 XP for future use
ℹtem Revised Total: 32 XP (bank 2 XP)
\end{itemize}

\section*{6. The Survivor} ∈dex{Character Examples!Survivor}

\textbf{Concept:} Someone who has endured hardship and developed resilience. Scars are maps; read them well.

\textbf{Typical Inspiration:} Veterans, refugees, outcasts, hardened adventurers

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Spirit, Body
ℹtem Skills: Endurance, Survival, (optionally) Perception/Insight
ℹtem Talents: Endurance, Adaptable
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Enduring difficult conditions
ℹtem Overcoming physical and mental challenges
ℹtem Using experience to avoid dangers
ℹtem Helping others survive hardships
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Improve physical and mental resilience
ℹtem Develop survival−related skills
ℹtem
```latex
Acquire better equipment and resources
ℹtem Learn to teach survival to others
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem What trauma or hardship have they survived?
ℹtem How has their past shaped their present?
ℹtem What survival skills have saved them repeatedly?
ℹtem How do they help others facing similar challenges?
\end{itemize}

\textbf{Build Blocks. Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes: Spirit 3 (9), Body 2 (6), Wits 1 (3), Presence 1 (3) → 21 XP
ℹtem Skills: Endurance 2 (4), Survival 2 (4) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Talent: Endurance (3 XP) using banked 1 + 4 = 5 XP; bank 2 XP
ℹtem Revised Total: 32 XP (bank 2 XP)
\end{itemize}

\section*{7. The Innovator} ∈dex{Character Examples!Innovator}

\textbf{Concept:} A creative problem−solver who finds new solutions. Blueprints on napkins, tomorrow in your pocket.

\textbf{Typical Inspiration:} Inventors, strategists, reformers, visionaries

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Wits, Presence
ℹtem Skills: Craft, Lore, Investigation
ℹtem Talents: Creative/Innovative thinking, Quick Study
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Finding novel solutions to problems
ℹtem Creating new devices or methods
ℹtem Analyzing systems for improvement
ℹtem Convincing others to try new approaches
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Develop specific technical specialties
ℹtem Create increasingly complex inventions
ℹtem Build support for innovative ideas
ℹtem Overcome resistance to change
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem What problem are they trying to solve?
ℹtem How do others react to their innovations?
ℹtem What unintended consequences might their creations have?
ℹtem Who benefits or suffers from their changes?
\end{itemize}

\textbf{Build Blocks. Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes: Wits 3 (9), Presence 2 (6), Body 1 (3), Spirit 1 (3) → 21 XP
ℹtem Skills: Craft 2 (4), Lore 2 (4) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Talent: Quick Study (3 XP) using banked 1 + 4 = 5 XP; bank 2 XP
ℹtem Revised Total: 32 XP (bank 2 XP)
\end{itemize}

\section*{8. The Networker} ∈dex{Character Examples!Networker}

\textbf{Concept:} Someone who builds and leverages social connections. A web of favors, a chorus of names.

\textbf{Typical Inspiration:} Merchants, spies, socialites, community leaders

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary: Presence, Wits
ℹtem Skills: Sway, Lore, (optionally) Command/Deception
ℹtem Talents: Network Builder, Command Presence / Silver Tongue
\end{itemize}

\textbf{Play Style:}
\begin{itemize}
ℹtem Building and maintaining relationships
ℹtem Gathering information through contacts
ℹtem Leveraging social influence
ℹtem Navigating complex social situations
\end{itemize}

\textbf{Development Path:}
\begin{itemize}
ℹtem Expand social network and influence
ℹtem Develop specific community ties
ℹtem Acquire political or economic power
ℹtem Master manipulation or leadership techniques
\end{itemize}

\textbf{Story Hooks:}
\begin{itemize}
ℹtem What networks or communities are they part of?
ℹtem How do they balance multiple relationships?
ℹtem What happens when loyalties conflict?
ℹtem How do they handle betrayal or broken trust?
\end{itemize}

\textbf{Build Blocks. Starting Build (32 XP).}
\begin{itemize}
ℹtem Attributes: Presence 3 (9), Wits 2 (6), Body 1 (3), Spirit 1 (3) → 21 XP
ℹtem Skills: Sway 2 (4), Lore 2 (4) → 8 XP
ℹtem Total: 29 XP (bank 1 XP)
\end{itemize}

\textbf{With Bonds/Complications (36XP).}
\begin{itemize}
ℹtem Add Talent: Silver Tongue (3 XP) using banked 1 + 4 = 5 XP; bank 2 XP
ℹtem Revised Total: 32 XP (bank 2 XP)
\end{itemize}

\section*{Creating Your Own Concept} ∈dex{Character Creation!original}

Start with Narrative
\begin{itemize}
ℹtem What is your character's background and motivation?
ℹtem What role do they play in their community or society?
ℹtem What relationships are important to them?
ℹtem What goals are they pursuing?
\end{itemize}

Add Mechanical Support
\begin{itemize}
ℹtem Choose attributes that support your concept
ℹtem Select skills that reflect their training and experience
ℹtem Consider talents that provide unique capabilities
ℹtem Think about assets that represent their resources
\end{itemize}

Consider Group Role
\begin{itemize}
ℹtem How does your concept complement other party members?
ℹtem What gaps in group capability can you fill?
ℹtem What unique contributions can you make?
ℹtem How will you work with other characters?
\end{itemize}

Plan for Growth
\begin{itemize}
ℹtem What short−term improvements make sense?
ℹtem What long−term development aligns with your concept?
ℹtem How might your character change over time?
ℹtem What legacy do you want to build?
\end{itemize}

\begin{tcolorbox}[colback=gray!5!white, colframe=gray!75!black, title=Character Concept Worksheet, fonttitle=\bfseries]
\textbf{Narrative Elements:}
\begin{itemize}
ℹtem Concept: ________________________________________
ℹtem Motivation: ________________________________________
ℹtem Background: ________________________________________
ℹtem Relationships: ________________________________________
\end{itemize}

\textbf{Mechanical Foundation:}
\begin{itemize}
ℹtem Primary Attributes: ________________________________________
ℹtem Key Skills: ________________________________________
ℹtem Starting Talents: ________________________________________
ℹtem Initial Assets: ________________________________________
\end{itemize}

\textbf{Development Plan:}
\begin{itemize}
ℹtem Short−term goals: ________________________________________
ℹtem Long−term vision: ________________________________________
\end{itemize}
\end{tcolorbox}

\section*{Final Notes} ∈dex{Character Creation!conclusion}

Remember that these examples are starting points, not limitations. The most interesting characters often combine elements from multiple concepts or create entirely new approaches. Work with your Game Master to ensure your character concept fits the campaign and provides engaging story opportunities.

The best characters are those that you find interesting to play and that contribute to an enjoyable experience for everyone at the table.

\chapter{World Regions and Cultures} łabel{ch:regions}

The world of Fate's Edge is a tapestry of ancient empires, emerging kingdoms, and untamed wilderness. This chapter surveys major regions and cultures that shape the setting−−−from the marble cities of Ecktoria to the mist−shrouded fields of Aelinnel. These frameworks are yours to adapt, blend, or reimagine.

\section*{The Amaranthine Inland Sea} ∈dex{Regions!Amaranthine Sea}

At the heart of the known world lies the Amaranthine Inland Sea, a wind−gnarled waterway ringed by marble quays, vineyard hills, and smoke−blue mountains. For millennia it has served as the circulatory system of trade, faith, and conquest. Tides are subtle, but seasonal winds and river−feeds set the rhythm of commerce, pilgrimage, and war.

\section*{Northern Shore of the Amaranthine Sea} ∈dex{Regions!Northern Shore}

\subsection*{Ecktoria −−− The Utaran Imperium Successor}

Once the furnace of empire (Marble & Fire), Ecktoria remains a palimpsest of power: old stones bearing new banners, old laws written under fresh seals. Though imperial reach waned, its civic habits endure.

\textbf{Marble Cities} Forums, amphitheaters, and aqueducts yet flow. District fountains double as public oaths guaranteed by guild charters.

\textbf{Imperial Roads} Mile−markers of white granite, way−shrines and customary tolls noted for couriers of the Ashen Staves.

\textbf{Legal Legacy} The Utaran Civic Codes govern contracts, inheritance, and war−rights; local custom bends them under licensed variance.

\textbf{Architectural Wonders} Sun−bridges spanning deltas, the Vault of a Thousand Maps, and the Amber Arch petrified by alchemical storm.

\subsection*{Acasia −−− "The Broken Province"}

Frontiers braided from roads, rivers, and resentments. Here the outer seams of empire frayed first. Fortresses turned manors, manors turned townholds, and banners multiplied like thistles after rain.

\textbf{Petty Kingdoms} Dozens of river−vales ruled by river−kings and banner−queens. Alliances shift with marriages, harvests, and omens.

\textbf{Fortified Towns} Walls for defense, not display. Gate−streets kink for ambush; towers carry horncodes every child knows.

\textbf{Mercenary Culture} Free companies keep a Black Ledger: contracts fulfilled, oaths kept, debts paid.

\textbf{Cultural Mix} Imperial rites meet clan feasts; old gods share niches with civic saints. Exiles and second chances (see Silkstrand tales) are common.

\subsection*{Vhasia −−− "Old Vhasia & The Bloodlands"}

Politically fractured land of courtly intrigue and martial tradition, where ancient bloodlines vie for supremacy amid shifting alliances and ceremonial warfare.

\textbf{Fortress Castles} Stone keeps crowned with gilded spires; courtiers plot in tapestried halls while knights train in courtyards.

\textbf{Political Intrigue} Complex web of alliances, vendettas, and ceremonial duels that settle matters of honor and succession.

\textbf{Court Culture} Elaborate ceremonies, patronage of arts, and rigid social hierarchies maintained through ritual and reputation.

\textbf{Heraldic Traditions} Complex system of banners, titles, and precedence that govern social interactions and military commands.

\subsection*{Thepyrgos}

Province and capital city renowned as a center of learning, magic, and scholarly pursuit, where ancient towers house both wisdom and dangerous secrets.

\textbf{Scholarly Traditions} Tower−cities where mages, philosophers, and researchers pursue knowledge in specialized colleges and scriptoriums.

\textbf{Arcane Heritage} Deep traditions of magical study, with libraries containing texts predating the fall of ancient empires.

\textbf{Academic Rivalries} Intense competition between schools of thought, often manifesting in formal debates, magical duels, or scholarly contests.

\textbf{Mystical Dangers} Forbidden knowledge and experimental magic that sometimes escape control, creating ongoing threats.

\subsection*{Viterra −−− "The Last Kingdom"}

Tudor−inspired realm that straddles the Dolmis and Amaranthine seas, known for its legalistic approach to governance and strategic river crossings.

\textbf{Hedge−Law Culture} Complex system of legal precedents, tolls, and river rights that govern everything from trade to personal conduct.

\textbf{Duchy System} Semi−autonomous regions governed by dukes who maintain their own courts and armies while owing fealty to the crown.

\textbf{River Commerce} Economy built around controlling strategic crossings, ferry rights, and maritime trade routes.

\textbf{Legalistic Politics} Intrigue centered on court cases, charter disputes, and the interpretation of ancient laws rather than open warfare.

\subsection*{Ubral −−− "The Stone Between Spears"}

Highland realm of rugged clans and fortified holds, where honor culture and martial traditions dominate social interactions.

\textbf{Clan Strongholds} Fortified positions in mountain passes and high valleys, each clan maintaining its own laws and customs.

\textbf{Honor Culture} Society built around concepts of personal honor, family reputation, and the resolution of disputes through formal challenges.

\textbf{Highland Warfare} Military traditions emphasizing heavy infantry, defensive positions, and knowledge of mountain terrain.

\textbf{Clan Loyalties} Complex web of alliances, blood−feuds, and marriage pacts that shift with each generation.

\subsection*{Kahfagia −−− "The Empire of Wakes and Storm−Flags"}

Maritime empire built on naval supremacy and exploration, where ship captains and merchant−adventurers shape both policy and culture.

\textbf{Naval Supremacy} Military and economic power based on controlling sea lanes, harbors, and maritime trade routes.

\textbf{Explorer Culture} Tradition of venturing into unknown waters, mapping new territories, and establishing trading posts.

\textbf{Storm−Flag Protocol} Complex system of maritime signals, weather prediction, and naval customs that govern seaborne activities.

\textbf{Mixed Heritage} Cosmopolitan society influenced by contacts with distant lands and diverse cultures encountered through exploration.

\section*{Southern Reaches} ∈dex{Regions!Southern}

\subsection*{Theona −−− "The Marsh Crown"}

Three island realms connected by causeways and maritime traditions, where wetland resources and naval culture define daily life.

\textbf{Marsh Agriculture} Sophisticated systems of dikes, canals, and floating gardens that support dense populations in wetland environments.

\textbf{Island Culture} Distinct traditions for each island, unified by shared maritime customs and inter−island trade.

\textbf{Waterborne Commerce} Economy based on fishing, water transport, and control of strategic waterways between islands.

\textbf{Folk Horror Traditions} Deep connection to marsh spirits, water deities, and ancient practices that blur the line between protection and appeasement.

\subsection*{The Mistlands −−− "Fields Under a Moving Sky"}

Isolated region shrouded in perpetual mists, where ancient Aelerian protectorate status creates tension between autonomy and oversight.

\textbf{Mistbound Geography} Landscape of bogs, waterways, and hidden settlements connected by causeways and boat paths.

\textbf{Bell Culture} Complex system of bells and wards must be maintained to keep the Direwood horrors at bay.

\textbf{Isolation Tensions} Cultural friction between desire for independence and practical need for trade and protection.

\textbf{Ancient Secrets} Ruins and artifacts predating the Aelerian protectorate, hinting at older civilizations and forgotten magics.

\section*{Peoples and Cultures} ∈dex{Cultures}

\subsection*{Wood Elves (Lethai−al "People of the Body")}

Inhabitants of the Valewood, deeply connected to the natural world and the cycles of growth and decay.

\textbf{Forest Harmony} Lifestyle integrated with woodland ecosystems, practicing sustainable hunting, gathering, and cultivation.

\textbf{Body−Centric Philosophy} Belief system emphasizing physical experience, instinct, and the wisdom of the body over abstract thought.

\textbf{Living Magic} Spellcasting traditions that work with natural forces rather than commanding them, often involving plant growth and animal communication.

\textbf{Seasonal Rituals} Calendar of ceremonies marking natural cycles, from planting rites to autumn harvests to winter hibernation periods.

\subsection*{High Elves (Lethai−thora "People of the Mind")}

Primarily found in Thepyrgos as established immigrants, known for their scholarly pursuits and intellectual traditions.

\textbf{Scholarly Excellence} Deep traditions of academic study, magical research, and philosophical debate.

\textbf{Mind−Centric Philosophy} Cultural emphasis on reason, logic, and the pursuit of abstract knowledge over physical concerns.

\textbf{Arcane Mastery} Advanced magical techniques and theoretical understanding that often surpass other traditions.

\textbf{Long Perspective} Tendency to view problems and conflicts through the lens of centuries or millennia rather than immediate concerns.

\subsection*{"Dark Elves" (Lethai−ar)}

Rare practitioners pledged to Isoka and Inaea, embracing serpent and spider themes without inherent evil, representing different philosophical approaches.

\textbf{Serpent Wisdom} Followers of Isoka, emphasizing transformation, renewal, and the shedding of old identities for new growth.

\textbf{Spider Webs} Devotees of Inaea, focusing on connections, patterns, and the weaving of fate through careful manipulation.

\textbf{Philosophical Balance} Neither inherently good nor evil, but representing alternative approaches to power and influence.

\textbf{Cultural Rarity} Uncommon in most settings, often viewed with suspicion or fascination by other cultures.

\subsection*{Gnomes (Aelinnel −−− "People of Sums")}

Inhabitants of the Mistlands, inspired by dark fairy tales and Wonderland lore, known for their mathematical precision and otherworldly logic.

\textbf{Mathematical Culture} Society built around complex calculations, probability, and the belief that all phenomena can be understood through numerical relationships.

\textbf{Fey Logic} Non−linear thinking patterns that seem illogical to outsiders but follow their own internal consistency.

\textbf{Mist Adaptation} Unique abilities to navigate and manipulate the perpetual mists of their homeland.

\textbf{Contract Culture} Deep tradition of precise agreements, wordplay, and the careful crafting of obligations that can be both helpful and dangerous.

\subsection*{Halflings (Aelaerem −−− "People of the Hearth")}

Peaceful agricultural communities with folk horror elements, emphasizing home, family, and the dark side of domestic tranquility.

\textbf{Hearth Culture} Deep connection to home, family, and the maintenance of traditional ways of life.

\textbf{Agricultural Expertise} Sophisticated farming techniques and seasonal celebrations that mark the rhythm of rural life.

\textbf{Folk Horror Elements} Dark undercurrents in seemingly peaceful communities, where hospitality can become trapping and tradition carries hidden costs.

\textbf{Community Defense} Strong traditions of mutual aid and collective action when the community is threatened.

\subsection*{Dwarves (Aeler −−− "People of Stone")}

Mountain−dwelling peoples of the Aelerian ranges, known for their craftsmanship, clan traditions, and complex relationship with stone and metal.

\textbf{Stone−Sense} Innate ability to understand and work with geological formations, making them master miners and architects.

\textbf{Clan System} Complex social structure based on family lines, with intricate systems of honor, debt, and mutual obligation.

\textbf{Craft Traditions} Legendary skills곘Ἥ

Looking at the partial document you provided, I can see it's a comprehensive RPG system called "Fate's Edge" that was cut off mid−sentence in Chapter 10. Let me complete this document by continuing from where it left off:

```latex
\documentclass[11pt,twoside,openany]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{titletoc}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage{makeidx}

≥ometry{letterpaper,margin=1in}
\makeindex

ℎypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
    pdftitle={Fate's Edge Player's Guide},
    pdfauthor={Nicholas A. Gasper}
}

\title{Fate's Edge: Player's Guide}
\author{Nicholas A. Gasper}
\date{October 12, 2025}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{þepage}
\fancyhead[LO]{≤ftmark}
\fancyhead[RE]{\rightmark}

\begin{document}

\maketitle

þispagestyle{empty}
\begin{center}
\textbf{Fate's Edge: Player's Guide}\\
\textcopyright\ Nicholas A. Gasper\\[1em]

This work is licensed under the Creative Commons Attribution 4.0 International License.\\
To view a copy of this license, visit:\\
\url{http://creativecommons.org/licenses/by/4.0/}
\end{center}

\tableofcontents

\chapter{World Regions and Cultures} łabel{ch:regions}

The world of Fate's Edge is a tapestry of ancient empires, emerging kingdoms, and untamed wilderness. This chapter surveys major regions and cultures that shape the setting−−−from the marble cities of Ecktoria to the mist−shrouded fields of Aelinnel. These frameworks are yours to adapt, blend, or reimagine.

\section*{The Amaranthine Inland Sea} ∈dex{Regions!Amaranthine Sea}

At the heart of the known world lies the Amaranthine Inland Sea, a wind−gnarled waterway ringed by marble quays, vineyard hills, and smoke−blue mountains. For millennia it has served as the circulatory system of trade, faith, and conquest. Tides are subtle, but seasonal winds and river−feeds set the rhythm of commerce, pilgrimage, and war.

\section*{Northern Shore of the Amaranthine Sea} ∈dex{Regions!Northern Shore}

\subsection*{Ecktoria −−− "The Utaran Imperium Successor"}

Once the furnace of empire (Marble & Fire), Ecktoria remains a palimpsest of power: old stones bearing new banners, old laws written under fresh seals. Though imperial reach waned, its civic habits endure.

\textbf{Marble Cities} Forums, amphitheaters, and aqueducts yet flow. District fountains double as public oaths guaranteed by guild charters.

\textbf{Imperial Roads} Mile−markers of white granite, way−shrines and customary tolls noted for couriers of the Ashen Staves.

\textbf{Legal Legacy} The Utaran Civic Codes govern contracts, inheritance, and war−rights; local custom bends them under licensed variance.

\textbf{Architectural Wonders} Sun−bridges spanning deltas, the Vault of a Thousand Maps, and the Amber Arch petrified by alchemical storm.

\subsection*{Acasia −−− "The Broken Province"}

Frontiers braided from roads, rivers, and resentments. Here the outer seams of empire frayed first. Fortresses turned manors, manors turned townholds, and banners multiplied like thistles after rain.

\textbf{Petty Kingdoms} Dozens of river−vales ruled by river−kings and banner−queens. Alliances shift with marriages, harvests, and omens.

\textbf{Fortified Towns} Walls for defense, not display. Gate−streets kink for ambush; towers carry horncodes every child knows.

\textbf{Mercenary Culture} Free companies keep a Black Ledger: contracts fulfilled, oaths kept, debts paid.

\textbf{Cultural Mix} Imperial rites meet clan feasts; old gods share niches with civic saints. Exiles and second chances (see Silkstrand tales) are common.

\subsection*{Vhasia −−− "Old Vhasia & The Bloodlands"}

Politically fractured land of courtly intrigue and martial tradition, where ancient bloodlines vie for supremacy amid shifting alliances and ceremonial warfare.

\textbf{Fortress Castles} Stone keeps crowned with gilded spires; courtiers plot in tapestried halls while knights train in courtyards.

\textbf{Political Intrigue} Complex web of alliances, vendettas, and ceremonial duels that settle matters of honor and succession.

\textbf{Court Culture} Elaborate ceremonies, patronage of arts, and rigid social hierarchies maintained through ritual and reputation.

\textbf{Heraldic Traditions} Complex system of banners, titles, and precedence that govern social interactions and military commands.

\subsection*{Thepyrgos}

Province and capital city renowned as a center of learning, magic, and scholarly pursuit, where ancient towers house both wisdom and dangerous secrets.

\textbf{Scholarly Traditions} Tower−cities where mages, philosophers, and researchers pursue knowledge in specialized colleges and scriptoriums.

\textbf{Arcane Heritage} Deep traditions of magical study, with libraries containing texts predating the fall of ancient empires.

\textbf{Academic Rivalries} Intense competition between schools of thought, often manifesting in formal debates, magical duels, or scholarly contests.

\textbf{Mystical Dangers} Forbidden knowledge and experimental magic that sometimes escape control, creating ongoing threats.

\subsection*{Viterra −−− "The Last Kingdom"}

Tudor−inspired realm that straddles the Dolmis and Amaranthine seas, known for its legalistic approach to 3
