% --- Patron: Raéyn, Keeper of the Sealed Gate (Thresholds & Warding) ---
\subsection{Raéyn, Keeper of the Sealed Gate (Thresholds \& Warding)}
\textit{Lore.} Raéyn is invoked at every border. His gift is the lock that preserves and the seal that keeps chaos at bay.

\begin{quote}
The door is not shut until Raéyn’s mark is traced.
\end{quote}

\paragraph{Seal the Latch (Low, 4 XP)} \emph{Action; Near; Yes (object).}
\textbf{Materials:} Trace a key-sign with ash/chalk.\\
\textbf{Effect:} Secure a container/door/gate. Attempts to open require intruder to generate \textbf{+1 SB} (Spades/Clubs).\\
\textbf{Push It:} First touch flares (noise/flash), revealing the attempt.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Quiet Gate (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A key turned backwards in a lock.\\
\textbf{Effect:} Warded threshold. Passing uninvited imposes \(-1\) die on the trespasser’s next action.\\
\textbf{Push It:} The ward whispers intruder’s purpose in one phrase.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Mark of Raéyn (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A drawn circle or sigil.\\
\textbf{Effect:} Protect a room/wagon. Crossing without consent generates \textbf{1 SB} (suit by GM).\\
\textbf{Push It:} Suppress one minor spell crossing; when it collapses, mark \textbf{1 SB (Hearts)} backlash.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Sealed Mouth (Standard, 7 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Thread tied across lips, then removed.\\
\textbf{Effect:} Choose one lost channel: speech, script, or gesture. Within, it fails for the scene.\\
\textbf{Push It:} Suppress all three, but you are muted until scene end.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Sealed Gate \textnormal{[WARD]} (High, 11 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Iron-powder circle, locked with a key.\\
\textbf{Effect:} Impassable boundary (~10 ft). Forcing entry inflicts \textbf{2 SB} (Clubs/Spades) on the intruder.\\
\textbf{Push It:} Enclose a chamber/courtyard. Each additional hour risks \textbf{1 SB (Diamonds)} toward omen/strain.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.
