% --- Patron: Khemesh, the Kraken Lord (Depths & Inevitable Power) ---
\subsection{Khemesh, the Kraken Lord (Depths \& Inevitable Power)}
\textit{Lore.} Khemesh is the unseen weight beneath the waves—the patience of the abyss and the certainty of drowning.

\begin{quote}
In the end, all things sink. The depths remember everything.
\end{quote}

\paragraph{Grasp of the Abyss (Low, 4 XP)} \emph{Action; Near; Yes (single target).}
\textbf{Materials:} A knot tied and dropped in water.\\
\textbf{Effect:} Phantom tentacles clutch; target suffers \(-1\) die next action or is briefly held in place.\\
\textbf{Push It:} Inflict 1 Harm (Crush); you gain Fatigue 1 (phantom drowning).\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Drowning Silence (Low, 5 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A seashell filled with water, then broken.\\
\textbf{Effect:} Muffle sound; rolls relying on voice/noise suffer \(-1\) die; Stealth gains +1 die.\\
\textbf{Push It:} Silence deepens to pressure; others resist or suffer Fatigue 1.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Weight of the Deep (Standard, 7 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A stone dropped into black water.\\
\textbf{Effect:} Target moves as if burdened; \(-1\) die to physical actions for the scene.\\
\textbf{Push It:} Briefly pin the target to deck/ground; mark \textbf{1 SB (Clubs)} from collateral strain.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Crushing Coil (Standard, 9 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} Rope/chain wrapped around your arm.\\
\textbf{Effect:} Spectral tentacle lashes or constricts: deal 2 Harm (Crush) to one target in Near.\\
\textbf{Push It:} Instead immobilize for one beat; you suffer Fatigue 1.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Abyssal Dominion (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Seawater poured in a circle.\\
\textbf{Effect:} Unseen tides hinder foes: enemies in Zone \(-1\) die to move/attack; allies \textbf{+1 die} to resist.\\
\textbf{Push It:} One massive tentacle strikes (3 Harm) once; draws the Keeper’s tide: \textbf{2 SB (Clubs/Spades)}.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.
