% --- Patron: The Clockwork Monad (Mechanism & Process) ---
\subsection{The Clockwork Monad (Mechanism \& Process)}
\textit{Lore.} Gears remember the plan even when their makers forget. The Monad prizes precision, iteration, and the beauty of mechanisms that keep their word.

\begin{quote}
Every tooth matters. Especially the ones you cannot see.
\end{quote}

\paragraph{Calibrated Touch (Low, 4 XP)} \emph{Action; Self; Yes (tinkering).}
\textbf{Materials:} A single oiled gear tooth.\\
\textbf{Effect:} \textbf{+1 die} to repair, set, or disarm precise mechanisms; re-roll one \texttt{1} on a Tinker action.\\
\textbf{Push It:} Guarantee no collateral on a simple device, but you take \textbf{1 SB (Clubs)} if rushed.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Process Lock (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A drop of red oil.\\
\textbf{Effect:} Name a process (reload, dispatch, raise alarm). First attempt to perform it in-scene suffers \(-1\) die as steps “stick.”\\
\textbf{Push It:} The second attempt also hesitates (one beat), but the third surges forward and \emph{forces 1 SB (Spades)}.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Iterative Advantage (Standard, 7 XP)} \emph{Action; Self/Ally; No.}
\textbf{Materials:} Notched tally on metal.\\
\textbf{Effect:} On a repeated action this scene, grant \textbf{+2 dice} or \textbf{+1 effect}.\\
\textbf{Push It:} Bank a follow-up \textbf{+1 die} for the same action later this scene; if unused, generate \textbf{1 SB (Diamonds)} as unused potential jams something.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Suppress Malfunction \textnormal{[DISPEL]} (Standard, 8 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} Mainspring coil released.\\
\textbf{Effect:} End/suppress an ongoing \emph{mechanical or procedural} failure/curse (e.g., jamming jam, recursive alarm, fatigue spiral tick). DV by fiction.\\
\textbf{Push It:} Also clear one related clock segment; create \textbf{1 SB (Clubs)} as pressure shifts elsewhere.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Prime the Engine (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A ring of interlocked cogs.\\
\textbf{Effect:} Allies in Zone add a “\emph{Process Buff}”: first action that repeats in-scene gets an automatic Position bump \emph{or} upgrades Effect.\\
\textbf{Push It:} Also seize timing: once, reorder two adjacent beats; mark \textbf{2 SB (Diamonds)} as the world protests.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.
