% --- Patron: Nidhoggr (Deep Earth & Rot) ---
\subsection{Nidhoggr (Deep Earth \& Rot)}
\textit{Lore.} What grows must fall. What stands must sink. Nidhoggr is patient ruin—roots splitting stone, rot reclaiming pride.

\begin{quote}
Lie down. The earth knows what to make of you.
\end{quote}

\paragraph{Root-Grip (Low, 4 XP)} \emph{Action; Near; Yes (grounded target).}
\textbf{Materials:} A seed pressed into soil.\\
\textbf{Effect:} Roots or sinkholes hinder a grounded foe; \(-1\) die on their next move/strike.\\
\textbf{Push It:} Brief immobilize (one beat) if they stand on soil/wood; you take \textbf{1 SB (Clubs)} as structures complain.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rot’s Kiss (Low, 5 XP)} \emph{Scene; Touch; No.}
\textbf{Materials:} A smear of compost or mold.\\
\textbf{Effect:} Tag gear or cover: first use this scene is \emph{Limited Effect} as rot softens it.\\
\textbf{Push It:} Also chip 1 integrity from mundane barriers/props touched in the scene.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Burden of Old Stone (Standard, 7 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A river-stone that has never seen sunlight.\\
\textbf{Effect:} Drop morale and vigor: \(-1\) die to strenuous actions for enemies who can feel the ground; allies gain steadiness (+1 die to resist knockdown).\\
\textbf{Push It:} Collapse a minor lintel/ledge to reshape cover; creates \textbf{1 SB (Clubs)}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Devour the Pillar (Standard, 9 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} Obsidian spindle with a hairline flaw.\\
\textbf{Effect:} Target a single support, axle, or keystone; reduce its integrity drastically (Keeper: equivalent to a big bite out of a [4] barrier).\\
\textbf{Push It:} Chain reaction threatens another support in Zone; you must choose to save someone or secure footing (\textbf{1 SB (Spades/Clubs)}).\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Grave-Quiet Dominion (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Fossil tooth shard and grave loam.\\
\textbf{Effect:} The ground asserts itself: enemies in Zone suffer \(-1\) die to rush/charge/leap; all falls worsen by one step; once, open a sink to isolate a foe.\\
\textbf{Push It:} Call the slow crush: \emph{force 2 SB (Clubs/Spades)} as supports, beams, or roots shift at the worst time.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.
