% =========================
% PATRON — The Carrion-King
% =========================
\subsection{The Carrion-King (Decay \& Cycles)}

\paragraph{Lore.}
The Carrion-King does not bring rot to destroy but to recycle. To him, decay is not an end but a return to the great cycle of flesh into soil, bone into earth. His priests are gardeners of entropy, ensuring that nothing lingers beyond its time and that new life always rises from death.

\paragraph{Quote.}
\emph{“All banquets end in bones, and from those bones the feast begins anew.” — The Carrion-King}

\paragraph{Rite of Gentle Rot (Low, 5 XP)} \emph{Instant; Touch; Yes (decay only).}
\textbf{Materials:} Spoiled food or a dead insect. \\
\textbf{Effect:} Accelerate natural decay on a small, non-living object (rope, lock, rations): +1 Effect to \emph{Break/Sabotage}. \\
\textbf{Push It:} A second, similar item in Close range decays; scavengers/vermin are drawn. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Wilting Bloom (Low, 4 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A withered flower. \\
\textbf{Effect:} Aura of mild decay: +1 die to resist disease/poison; your carried food won’t worsen (becomes tasteless). \\
\textbf{Push It:} Wither a small plant/food source in Near; suffer \textbf{Fatigue 1}. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Cycle's Turn (Standard, 8 XP)} \emph{Scene; Touch; No.}
\textbf{Materials:} A creature dead less than an hour. \\
\textbf{Effect:} From death, sustain life: choose one—purify a small food/water cache; sprout useful fungi/herbs; \emph{or} grant +1 die to resist disease/poison to one target. \\
\textbf{Push It:} Superior yield/quality; local life/death balance tilts (GM clocks/complications). \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Peaceful Rest (Standard, 7 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} Grave dirt over the corpse. \\
\textbf{Effect:} Lay a minor spirit; prevent easy animation \emph{or} quiet a small haunting. Gain +2 dice on the next social roll with mourners/spirits. \\
\textbf{Push It:} You/ally gain undead-fear immunity for the scene; other nearby spirits grow restless. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Final Compost (High, 13 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A handful of grave dirt. \\
\textbf{Effect:} Accelerate decay across a zone: structures falter (attackers relying on them $-1$ die), enemy maintenance falters ($-1$ die to keep gear/efforts stable). Alternatively, consume a major obstacle over the scene. \\
\textbf{Push It:} The zone erupts with sickly growth; creatures lingering gain \emph{Sickened}. The growth may be harvested later. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Great Cycle (High, 14 XP)} \emph{Extended; Touch; No.}
\textbf{Materials:} Bury a seed in rich, rotten earth. \\
\textbf{Effect:} Transform a significant dead mass (large corpse/fallen tree) into something useful: fertile plot, unique reagent, or a temporary environmental boon (GM scales time/impact). \\
\textbf{Push It:} Compress to one scene; spectacle attracts attention and interference. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.
