% =========================
% PATRON — The Gallow's Bell
% =========================
\subsection{The Gallow's Bell (Consequences \& Retribution)}

\paragraph{Lore.}
The Bell tolls for every oath broken, every sin unatoned. Its sound is not heard by all, but by those whose guilt cannot be outrun. To follow the Gallow’s Bell is to be an instrument of consequence, a reminder that no action escapes the echo of its price.

\paragraph{Quote.}
\emph{“Three times shall the bell toll, and thrice shall the guilty stumble.” — The Gallow’s Bell}

\paragraph{Rite of the Whispered Name (Low, 4 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Noose fragment; whispered confession. \\
\textbf{Effect:} Mark a target by name; within 3 scenes the GM must introduce a complication tied to their recent deeds. \\
\textbf{Push It:} The complication strikes immediately; homonyms may also be brushed by the omen. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Curse of the Unsettled Sleep (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Dream-catcher woven with guilt; midnight vigil. \\
\textbf{Effect:} Target dreams of misdeeds, inviting roleplay leverage and soft penalties per fiction. \\
\textbf{Push It:} The dream grants a true omen of coming consequence; one of your secrets is revealed to the dreamer. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Broken Mirror (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Shattered mirror; drop of the target’s blood. \\
\textbf{Effect:} Target’s reflection shows “true nature” to observers; their deceptions generate automatic narrative complications on success. \\
\textbf{Push It:} The reflection turns hostile (ambient social friction); all mirrors nearby become \emph{strange} until scene end. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Mark of Unfinished Business (Standard, 7 XP)} \emph{Extended; Touch; No.}
\textbf{Materials:} An unfinished letter; a broken chain link. \\
\textbf{Effect:} Target accrues story-weight from unresolved obligations; expect recurring complications until addressed. \\
\textbf{Push It:} The obligation focuses into a specific demand; you share narrative responsibility to resolve it. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Curse of the Singing Chain (High, 13 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A prison chain and a funeral bell-hammer. \\
\textbf{Effect:} Target hears echoes of broken promises: suffer \textbf{Fatigue 1}; lying draws narrative backlash. \\
\textbf{Push It:} A spectral chain manifests (movement constrained by fiction); nearby chains resonate \& misbehave. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{The Bell That Rings Thrice (High, 14 XP)} \emph{Extended; Zone; No.}
\textbf{Materials:} Gallows rope; three iron bells rung in order. \\
\textbf{Effect:} Over the next session, the target suffers three escalating, nature-true consequences (GM stages). \\
\textbf{Push It:} The toll splashes onto their close allies/kin; you are marked as an agent of retribution. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.
