# % --- Patron: The Carrion-King (Decay, Renewal & Transformation) ---

\subsubsection{The Carrion-King (Decay, Renewal \& Transformation)}
\textit{Lore.} The Carrion-King is the master of endings that become beginnings. He does not destroy, but transforms—turning death into new life, decay into opportunity, and endings into fresh starts. His followers are harvesters of potential, seeing in every fall the seeds of future growth.

\begin{quote}
What crumbles feeds what grows. What dies becomes the soil of tomorrow's triumph.
\end{quote}

\paragraph*{Rite of Consuming Rot (Low, 5 XP)} \emph{Instant; Touch; Yes (decay only).}
\textbf{Materials:} Organic matter in early stages of decay. \\
\textbf{Effect:} Accelerate natural decay to weaken or destroy: +2 Effect to \emph{Break/Sabotage} on organic materials (ropes, leather, wood). Gain 1 Boon if the decay creates an opportunity for you or allies. \\
\textbf{Invoke:} 1 action; mark +1 Obligation. \\
\textbf{Push It:} Spread decay to similar materials in Close range; mark 1 SB (Clubs) as the rot becomes noticeable. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph*{Rite of the Harvested End (Low, 4 XP)} \emph{Scene; Touch; No.}
\textbf{Materials:} The remains of a recently ended thing (burnt letter, wilted flower, shattered glass). \\
\textbf{Effect:} Extract value from endings: from a defeated enemy, gain +1 die to next action; from a failed plan, re-roll one 1 on your next roll; from a broken item, gain 1 SB to spend immediately. \\
\textbf{Invoke:} 1 action; mark +1 Obligation. \\
\textbf{Push It:} Harvest additional value but mark Fatigue 1 from dwelling on endings. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Fertile Death (Standard, 8 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Ashes, compost, or the remains of anything that once lived. \\
\textbf{Effect:} Transform death into growth: create beneficial terrain (cover, concealment, or advantageous positioning) OR grant allies +1 die to healing/recovery rolls. Choose one effect per scene. \\
\textbf{Push It:} Both effects apply but attract unwanted attention (vermin, scavengers, or curious onlookers). \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Transformed Spirit (Standard, 7 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} A token from a deceased being (hair, nail, written name). \\
\textbf{Effect:} Channel the essence of what was: gain one skill die reflecting the deceased's expertise for one scene OR ask one question about their knowledge/abilities. \\
\textbf{Push It:} The spirit's influence lingers - gain permanent insight (+1 die specialty) but suffer occasional possession-like effects (GM discretion). \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Great Consumption (High, 13 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A significant amount of organic matter (corpse, fallen tree, collapsed building). \\
\textbf{Effect:} Transform a large area through decay and renewal: choose two - create difficult terrain that favors you, summon Cap 3 swarm of scavengers as temporary allies, or generate valuable reagents worth 2 XP. \\
\textbf{Push It:} All three effects occur but start a 6-segment \textbf{Ecosystem Disruption} clock that will cause problems later. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Eternal Cycle (High, 14 XP)} \emph{Extended; Touch; No.}
\textbf{Materials:} The complete remains of something significant that has ended. \\
\textbf{Effect:} Complete a transformation cycle: destroy one major asset/enemy/obstacle and create something new of equal or greater value. GM and player collaborate to define the transformation. \\
\textbf{Push It:} The transformation is immediate and spectacular but creates a 6-segment \textbf{Cycle Debt} clock - the King will demand another significant ending soon. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\subsection*{Carrion-King's Corruption Table}
\label{sec:carrion-king-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Carrion's Insight: +1 die to Notice decay or hidden weaknesses in structures or beings. & Must inspect decay firsthand; suffer 1 Fatigue when exposed to fresh death or rot. \\
\midrule
2 & Deathward Sense: Once per session, detect the last living moment of a dead being within Close range. & Cannot lie about death you’ve witnessed; must correct falsehoods. \\
\midrule
3 & Rotblood Resilience: Gain +1 die to resist disease and poison. & Immune system adapts slowly; each new disease/poison requires 1 Fatigue to resist. \\
\midrule
4 & Glean from Grief: Once per scene, gain +1 die after witnessing a significant loss or defeat. & Compelled to linger at scenes of death; must spend one beat observing or risk 1 SB (Clubs). \\
\midrule
5 & Cycle's Whisper: You can sense the “next ending” in any process—ask the Keeper one question about how a situation will collapse or conclude. & Must speak the truth about what you see, even if it harms your position. \\
\midrule
6+ & Eternal Bloom: Once per session, declare a “death that births life.” Sacrifice an asset or ally to create something new of equal or greater value. & Mark +2 Obligation when using this power. \\
\bottomrule
\end{longtable}