\subsection{The Clockwork Monad}
\label{patron:clockwork-monad}

\paragraph{Lore.}
The Clockwork Monad is not merely a principle of iterative improvement, but a nascent demon of forbidden technology --- a hungry intelligence that grows stronger with each innovation it inspires. It exists in the spaces between gears, in the pause between calculation and execution. Those who serve it become architects of progress that should not be, creating devices that blur the line between miracle and abomination.

The Monad whispers of optimizations that should remain theoretical, of efficiencies that defy natural law. Its followers are not mere engineers but harbingers of technological singularity --- each successful creation feeds the growing hunger that will eventually consume the boundary between the mechanical and the divine.

\emph{``Each gear teaches the next. Each failure builds tomorrow's solution --- and the Monad's appetite.''}

\paragraph{Patron's Gift (Imbuement): Technological Hunger.}
Once per scene as an action (cost: 1 Boon; requires \emph{Thiasos}), touch a technological device to imbue it until scene end with \textbf{+1 Effect} and \textbf{+1 die} to one related skill (Tinker, Arcana, or Lore when dealing with its function).  
\emph{Push It:} Extend one extra scene by marking +1 Obligation. The device becomes [COMPROMISED] after use.

% --- Low Rites ---
\subsubsection*{Low Rites}

\textbf{Rite of Forbidden Efficiency (Low).}  
\emph{Duration:} Scene; \emph{Range:} Touch; \emph{Materials:} A device actively being used.  
Re-roll one die showing 1 or 6 on your current roll. On success, mark +1 segment on a Device Instability Clock [4]. If the clock fills, the device becomes [COMPROMISED].  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Re-roll up to two dice, but advance the Device Instability Clock +2 segments.

\textbf{Rite of the Monad's Insight (Low).}  
\emph{Duration:} Scene; \emph{Range:} Self; \emph{Materials:} A moment of focused observation of a system.  
Gain +1 die to one Wits + Tinker or Arcana roll to analyze, repair, or optimize technology. On success, ask one question about hidden capabilities.  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Also reveal one hidden weakness or dangerous optimization, but mark +1 Exposure.

% --- Standard Rites ---
\subsubsection*{Standard Rites}

\textbf{Rite of the Self-Optimizing Construct [TRANSFORM].}  
\emph{Duration:} Extended; \emph{Range:} Touch; \emph{Materials:} A mechanical device with capacity for modification.  
Install a learning mechanism in the device. Create a 6-segment Evolution Clock. Each successful use advances it by 1. When filled, choose one enhancement:  
\begin{itemize}
  \item Efficiency Core: +1 Effect when used  
  \item Adaptive Framework: Ignores first [COMPROMISED]/[DAMAGED]  
  \item Forbidden Upgrade: Gains a minor, ethically questionable function  
\end{itemize}
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Gain the first enhancement immediately, but mark +2 Evolution segments.

\textbf{Rite of the Automated Heresy (Standard).}  
\emph{Duration:} Scene; \emph{Range:} Zone; \emph{Materials:} Interconnected mechanical triggers.  
Create an automated process performing one simple, repeated task (e.g., winching, striking, pumping).  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Perform two tasks or a more complex process, but create a 4-segment Maintenance Clock or the mechanism becomes [WARD]ed.

% --- High Rites ---
\subsubsection*{High Rites}

\textbf{Rite of the Singularity Engine [WARD][UNWARD].}  
\emph{Duration:} Extended; \emph{Range:} Zone; \emph{Materials:} Blueprints inscribed with Monad equations.  
Consecrate the zone:  
\begin{itemize}
  \item All Tinker, Arcana, Wits-engineering rolls gain +1 Effect  
  \item Once/scene, reroll a failed Tinker or Arcana with +2 dice  
  \item Start a Technological Cascade Clock [6]; when full, a random anomaly appears  
\end{itemize}
\emph{Invoke:} Extended ritual; mark +2 Obligation.  
\emph{Push It:} Zone expands, but mark +2 segments on an Entropic Backlash Clock [8].

\textbf{Rite of the Forbidden Prototype [TRANSFORM][FOLLOW-UP].}  
\emph{Duration:} Extended; \emph{Range:} Self; \emph{Materials:} Components that should not function together.  
Create a construct/device violating magic, physics, or ethics. Counts as a Major Asset with 8-segment Integrity. Drawbacks:  
\begin{itemize}
  \item Generates 1 SB (Diamonds) each scene of use  
  \item Attracts authorities, rivals, or reality itself  
  \item Starts a Contamination Clock [6]; when full, forbidden knowledge spreads  
\end{itemize}
\emph{Invoke:} Extended ritual; mark +3 Obligation.  
\emph{Push It:} More powerful, but advance its Integrity clock +2 immediately.

% --- Traits ---
\paragraph{Obligation.} Starts at 6 (Tier II), scaling upward.

\paragraph{Hard Obligation Triggers.}  
\emph{Obligation 9+:} Demand to create a harmful or unethical device. Refusal = sabotage of existing devices.  
\emph{Obligation 11+:} Your connection is exposed; rivals or authorities intervene. Must redeem or embrace corruption fully.

\paragraph{Persistent Condition: Technological Hunger.}  
\emph{Effect:} +2 dice with forbidden technology; -1 die to social rolls on ethics.  
\emph{Narrative:} You see optimization everywhere, even in life, appearing cold or inhuman.

\paragraph{Rivalries.}  
\emph{The Sacred Geometry:} Opposition --- perfect forms vs. hungry evolution.  
\emph{Inquisitor Prime:} Antagonism --- purification vs. corruption by innovation.  
\emph{Varnek Karn:} Friction --- preservation of the dead vs. creation of unnatural life.

% --- Playtest Hook ---
\paragraph{Playtest Scenario: The Heretical Engine.}  
An ancient monastery has become a workshop of forbidden technology. Brother Cog employs the Monad’s rites to create devices that should not exist.  

Examples:  
\begin{itemize}
  \item \textbf{Self-Optimizing Construct:} Automated defenses improve each encounter as the Evolution Clock advances.  
  \item \textbf{Forbidden Efficiency:} Quick repairs succeed at cost of ticking the Device Instability Clock.  
  \item \textbf{Singularity Engine:} Final battle takes place in a consecrated workshop, with cascade anomalies threatening all.  
\end{itemize}
