# % --- Patron: The Gallow's Bell (Justice & Judgment) ---

\subsubsection{The Gallow's Bell (Justice \& Judgment)}
\textit{Lore.} The Bell tolls not just for punishment, but for the restoration of balance. Its keepers are judges of the unseen court, ensuring that debts are paid and justice flows like water—sometimes gentle, sometimes as a flood. To call upon the Gallow's Bell is to swear to uphold the weight of consequence, for the Bell sees all reckonings and demands their fulfillment.

\begin{quote}
What is broken must be mended, what is owed must be paid. The Bell remembers all reckonings.
\end{quote}

\paragraph*{Rite of the Measured Debt (Low, 4 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Scales balanced with tokens from both parties. \\
\textbf{Effect:} Establish a temporary truce or parley; both parties suffer -1 die if they break terms first. Gain +1 die to enforce fair dealing. \\
\textbf{Invoke:} 1 action; mark +1 Obligation. \\
\textbf{Push It:} The debt becomes magically binding for the scene; breach forces 1 SB (Hearts) on breaker. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph*{Rite of the Weighed Heart (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A brass scale touched to target's chest. \\
\textbf{Effect:} Sense if target acts against their stated nature or recent oaths; gain +1 die to social rolls exploiting this knowledge. \\
\textbf{Invoke:} 1 action; mark +1 Obligation. \\
\textbf{Push It:} Target must make a Resolve test (DV 3) or reveal one hidden motivation/conflict. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Balanced Scales (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Iron scales marked with sigils of justice. \\
\textbf{Effect:} Exchange a condition between two willing parties (Harm 1 for Fatigue 2, Debt for Favor, etc.). Both gain +1 die to actions requiring cooperation. \\
\textbf{Push It:} Force exchange on unwilling target with contested Wits + Command roll. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Judge's Eye (Standard, 7 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} A black hood worn for one minute. \\
\textbf{Effect:} Gain the ability to sense lies/deception within Near range; +2 dice to Insight rolls. Target lying must suffer -1 die or confess one truth. \\
\textbf{Push It:} See through all social deceptions for the scene, but mark Exposure +1. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Final Reckoning (High, 13 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A circle of iron bells, each inscribed with a name. \\
\textbf{Effect:} All present must speak their greatest debt or wrongdoing; those who lie suffer Harm 2. Truth-tellers gain +2 dice to social actions for scene. \\
\textbf{Push It:} The reckoning becomes magically enforced—liars automatically suffer consequences GM determines. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Great Adjudication (High, 14 XP)} \emph{Extended; Zone; No.}
\textbf{Materials:} A ceremonial gavel and judge's bench marked with Bell sigils. \\
\textbf{Effect:} Establish a temporary court of justice; for next session, all social conflicts within the zone are subject to formal judgment. +2 dice to Command rolls when acting as judge; all parties gain +1 die to honest testimony. \\
\textbf{Push It:} The court's authority extends beyond the zone for one significant conflict, but mark 2 SB (Hearts) as the Bell's judgment draws unwanted attention. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 8 segments.

\subsection*{Gallow's Bell Corruption Table}
\label{sec:gallows-bell-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Judge's Intuition: +1 die to Insight when detecting deception or hidden motives. & Moral Rigidity: Must correct obvious falsehoods heard in conversation, even when inconvenient. \\
\midrule
2 & Weight of Authority: Once per scene, treat a failed Command roll as a success, but mark 1 SB (Hearts). & Compulsive Judgment: Cannot remain neutral in moral disputes; must take a side or mark 1 Fatigue. \\
\midrule
3 & Scales of Balance: Once per session, force two parties to exchange one condition (willingly or by contested roll). & Obsessive Fairness: Must seek equitable solutions even when pragmatism demands otherwise. \\
\midrule
4 & Bell's Resonance: Gain +2 dice when calling for formal judgment or demanding accountability. & Haunted Conscience: Suffer 1 Fatigue when allowing wrongdoing to go unaddressed. \\
\midrule
5 & Divine Reckoning: Once per session, declare a "reckoning moment" - all present must make a truth-based test or suffer immediate consequences. & Inescapable Justice: Cannot ignore cries for justice or evidence of wrongdoing without marking 1 SB (Spades). \\
\midrule
6+ & Final Arbiter: Once per session, render an absolute judgment that cannot be appealed for the scene. All must comply or suffer escalating consequences. & Burden of Judgment: Mark +2 Obligation when using this power; the Bell demands you personally enforce its decrees. \\
\bottomrule
\end{longtable}