% --- Patron: Grimmir, the Old Man of the Forest (Instinct & Wild Memory) ---

\subsubsection{Grimmir, the Old Man of the Forest (Instinct \& Wild Memory)}
\textit{Lore.} Grimmir is not worshipped so much as remembered—the antlered shadow at the fire’s edge, the whisper in wolf-song, the hand that first showed humankind flint and flame. He is the wild within memory, neither wholly man nor wholly beast, the primeval pact that civilization dare not wholly forget. Those who call upon him gain keen senses, primal endurance, and the feral cunning of predator and prey alike—but risk leaving behind the tongue of men for the howl of beasts.

\begin{quote}
“They say he taught the wolves to stalk, and the children of men to kindle fire. Listen close, and you may remember what your blood already knows.”  
\end{quote}

\paragraph*{Hunter’s Snarl (Low, 4 XP)} \emph{Scene; Self; Yes.}\\
\textbf{Materials:} A tooth, claw, or shard of bone.\\
\textbf{Effect:} Gain +1 die to intimidate or resist fear by baring primal menace.\\
\textbf{Push It:} First foe to strike hesitates; their roll suffers -1 die.\\
\emph{Requires: Familiar.}

\paragraph*{Wolf-Step (Low, 5 XP)} \emph{Scene; Self; Yes.}\\
\textbf{Materials:} Soil from a predator’s track.\\
\textbf{Effect:} Gain +1 die to Stealth or Athletics when stalking, flanking, or slipping pursuit.\\
\textbf{Push It:} Ignore one minor terrain penalty (roots, brambles, mud).\\
\emph{Requires: Familiar.}

\paragraph{Horn of the Wild Hunt (Standard, 8 XP)} \emph{Scene; Near; Yes.}\\
\textbf{Materials:} A hollowed bone or carved horn.\\
\textbf{Effect:} Call the Hunt. Allies gain +1 die on next strike or chase; lesser foes must test Resolve or falter.\\
\textbf{Push It:} Summon a Cap 1 beast-ally (hound, raven, stag) until scene end.\\
\emph{Requires: Familiar + Codex.}

\paragraph{Mark of Antlers (Standard, 8 XP)} \emph{Scene; Self or Touch; Yes.}\\
\textbf{Materials:} Ash, ochre, or blood marked on brow.\\
\textbf{Effect:} Bear antler-crown. Gain +1 die to Command, resist shoves, and impose presence as larger-than-life.\\
\textbf{Push It:} For one exchange, foes in Near range must resist or strike at -1 die.\\
\emph{Requires: Familiar + Codex.}

\paragraph{Wild Communion (Major, 12 XP)} \emph{Scene; Region; Yes.}\\
\textbf{Materials:} Sit upon earth within a ring of bones or roots.\\
\textbf{Effect:} Ask two questions of beasts or primal memory concerning hunts, movements, or omens. Gain advantage to track or ambush.\\
\textbf{Push It:} Ask a third, or extend insight to human hunters and wildfolk.\\
\emph{Requires: Familiar + Codex.}

\paragraph{Shape of the Ancestor (Epic, 16 XP)} \emph{Scene; Self; Yes.}\\
\textbf{Materials:} Don a mask of fur, antler, or bone.\\
\textbf{Effect:} Transform into hybrid form (fangs, claws, heavy pelt). For the scene:  
\begin{itemize}
  \item +1 Body  
  \item Natural melee attack (+1 enchanted)  
  \item +1 die to Athletics, Melee, or Survival  
\end{itemize}  
\textbf{Push It:} Extend blessing to one ally marked in blood/ash.\\
\emph{Requires: Familiar + Codex + Tier III.}\\
\emph{Obligation:} 8 segments.

\subsection*{Grimmir’s Corruption Table}
\label{sec:grimmir-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Primal Senses: +1 die to Notice in wild terrain or when ambushed. & Bestial Quirk: Growl, sniff, or bare teeth when startled or angered. \\
\midrule
2 & Predator’s Instinct: Once per scene, gain +1 die when stalking or flanking. & Hunger Urge: Smell of blood or meat causes distraction; resist or mark Fatigue. \\
\midrule
3 & Wild Endurance: +1 die to resist fear, charm, or supernatural awe. & Feral Impulse: In stress, must test Resolve (DV 3) or lash out instinctively. \\
\midrule
4 & Pack Hunter: Allies in Near range gain +1 die when following your lead in combat. & Territorial: -1 die when far from wild places for more than a day. \\
\midrule
5 & Ancestral Memory: Once/session, recall instinctive skill or lost craft. & Speech Slips: Under stress, lapse into snarls, growls, or fragmented primal tongue. \\
\midrule
6+ & Alpha Presence: Once/session, command all beasts in Near range without roll. & Beastbound: Using this power invites the Ancestor—GM may impose primal compulsions unless resisted at cost of 1 SB (Clubs). \\
\bottomrule
\end{longtable}