% --- Patron: Ikasha, She Who Sleeps (Latent Potential & Shadow) ---
\subsubsection{Ikasha, She Who Sleeps (Latent Potential \& Shadow)}
\textit{Lore.} Ikasha is the hush between footfalls, the patience of dark water, the black-feathered watcher at every threshold. In stillness she gathers what might be, in crossroads she whispers of what may yet come. Ravens circle her, bearing secrets between worlds. Her followers learn to move unseen and speak unremembered, becoming shadows that slip between what is and what could be.

\begin{quote}
Blow out the candle. If the room listens back, ask softly. At the next crossroads, the raven waits—and the shadow remembers your passing.
\end{quote}

\paragraph{Touch the Umbral Veil (Low, 4 XP)} \emph{Action; Self; Yes (Stealth).}
\textbf{Materials:} A piece of black cloth.\\
\textbf{Effect:} Start \emph{Controlled} on one Stealth roll or gain +1 effect to hide/move quietly.\\
\textbf{Push It:} Brief shadow-muffling (ignore one noisy tell), but leave a shadow-double that may echo you later at an ill moment—mark \textbf{1 SB (Diamonds)} as the shadow takes on its own agenda.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Crossroads Raven (Low, 5 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Scatter three black feathers or carve a crossroads sign.\\
\textbf{Effect:} Summon an omen-raven; grant \textbf{+1 die} to a navigation, pursuit, or diversion action \emph{or} force an enemy to hesitate at a fateful moment.\\
\textbf{Push It:} The raven speaks one cryptic truth, but demands a secret in return—the raven's insight costs you \textbf{1 SB (Hearts)} as it glimpses something you'd rather keep hidden.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Draw from the Umbral Reservoir (Standard, 8 XP)} \emph{Action; Self/Ally; No.}
\textbf{Materials:} A vial of moonless-night water.\\
\textbf{Effect:} \textbf{+2 dice} to stealth, deception, or resolve \emph{or} clear \emph{Fatigue 1}.\\
\textbf{Push It:} Also gain one free escape attempt; next scene, you must help another cross a threshold or flee danger—shadow-debt calls for shadow-service.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Secret Keeper's Burden (Standard, 9 XP)} \emph{Instant; Touch; No.}
\textbf{Materials:} A lock of hair or intimate token.\\
\textbf{Effect:} Compel a truthful answer to one direct question (deep secrets may allow a Resolve test to resist).\\
\textbf{Push It:} Learn the answer \emph{and} a key hidden emotion; target learns one of your secrets in return, carried by a raven to them in dreams—mark \textbf{1 SB (Hearts)} as the secret exchange creates unexpected complications.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Become the Shadow at the Crossroads (High, 12 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Stand in absolute darkness or at a deserted crossroads.\\
\textbf{Effect:} Intangible to mundane harm; pass through thresholds and small gaps; \textbf{+2 dice} to Stealth; auto-succeed one escape. Cannot manipulate normal objects.\\
\textbf{Push It:} Interact once with a bound or thresholded object (a door, a lock, a sealed letter), but you become partially corporeal and vulnerable for one beat. Ravens may mark you—shadow-form destabilizes, mark \textbf{1 SB (Spades)} as reality asserts its claim.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.