# % --- Patron: Inaea, Angel of the Spider (Webs & Fate) ---

\subsubsection{Inaea, Angel of the Spider (Webs \& Fate)}
\textit{Lore.} Where Isoka sheds, Inaea binds—threads of debt, favor, and inevitability. She is the Weaver of Connections, the patron of those who see the invisible threads that tie fate to fate, person to person, promise to consequence. Her followers learn to manipulate these connections, drawing power from the web of relationships that binds all things.

\begin{quote}
Tie one knot for what you owe, two for what you're owed, and a third for what will answer both. But beware—the Weaver always collects her due.
\end{quote}

\paragraph*{Tie a Simple Knot (Low, 4 XP)} \emph{Action; Near; Yes (link once).}
\textbf{Materials:} A single thread.\\
\textbf{Effect:} Declare two minor events linked; either \textbf{force 1 SB} (GM suit) on a foe when the first triggers \emph{or} bank \textbf{+1 die} for a follow-on roll this scene.\\
\textbf{Invoke:} 1 action; mark +1 Obligation.\\
\textbf{Push It:} The held +1 ignores one minor disruption; the web's tension tightens—mark \textbf{1 SB (Clubs)} as the connection becomes more demanding.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph*{Rite of the Tangled Thread (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Tug a web or net.\\
\textbf{Effect:} Invisible snare in a lane/door. First to cross suffers \(-1\) die on next action.\\
\textbf{Invoke:} 1 action; mark +1 Obligation.\\
\textbf{Push It:} Brief bind (one beat) enabling an ally setup; affects all who cross—the tangled threads ensnare indiscriminately, mark \textbf{1 SB (Spades)} as allies may also be caught.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Weave the Strand of Inevitability (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Three colored threads woven.\\
\textbf{Effect:} Link two actors/actions: when A moves, B is exposed. Choose: \textbf{force 1 SB on B} next action \emph{or} \textbf{+2 dice} to one prediction/setup keyed to the link.\\
\textbf{Push It:} Invert once (B cues A); the web's pattern shifts—mark \textbf{1 SB (Hearts)} as the manipulation strains the natural order.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Weaver's Glance (Standard, 7 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Watch a spider finish one radial line.\\
\textbf{Effect:} Ask one precise question about in-scene ties; then gain \textbf{+1 effect} on one leverage/pressure action exploiting it.\\
\textbf{Push It:} Surface a hidden tie (Keeper reveals a quiet obligation/fear); mark \emph{Exposure +1} as the web exposes your own entanglements.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Bind the Bargain \textnormal{[OATH]} (High, 11 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Silk loop tied around two thumbs, then cut/knotted.\\
\textbf{Effect:} Bind up to two consenting parties to a clear term. Breach \emph{forces 2 SB} on the breaker and leaves a subtle tell until amends.\\
\textbf{Push It:} Widen to a small circle (up to four); each party names a narrow loophole (Keeper approves). Exploiting it generates \textbf{1 SB (Diamonds)} as the web's complexity creates unforeseen resonances.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.

\subsection*{Inaea's Corruption Table}
\label{sec:inaea-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Gentle Guidance: +1 die to Persuade when offering comfort or aid to those in distress. & Comfort Addiction: Must offer solace to those in pain when encountered, or mark 1 SB (Hearts). \\
\midrule
2 & Web Sense: Once per scene, sense the emotional connections between any two people within Near range. & Empathic Overload: Suffer 1 Fatigue when exposed to intense emotions or conflicts between others. \\
\midrule
3 & Binding Touch: Once per session, make a promise so compelling that the target must make a Resolve test (DV 4) or accept it unquestioningly. & Compulsive Caretaking: Feel responsible for the wellbeing of those you've aided; suffer -1 die when ignoring their needs. \\
\midrule
4 & Fate's Thread: Gain +2 dice when predicting how a relationship or alliance will develop. & Manipulative Instincts: When offering help, secretly desire control over the recipient's choices; mark 1 SB (Diamonds) when acting purely selflessly. \\
\midrule
5 & Spider's Mercy: Once per session, completely sever an unwanted emotional bond or harmful relationship for yourself or another. & Isolation Hunger: After severing connections, suffer 1 Fatigue and crave new bonds to fill the void. \\
\midrule
6+ & Weaver's Dominion: Once per session, declare that all relationships within a scene are subject to your subtle influence—you gain +2 dice to social manipulation, but the web tightens around all involved. & Threads of Control: Mark +2 Obligation when using this power; those affected become subtly dependent on your approval. \\
\bottomrule
\end{longtable}