\section{Of Patrons, Runes, and Invokers}
\label{sec:invoker-lore}

\begin{quote}
``You wish to walk the road of power? Then listen well. The world is old, and older still are the voices beneath it. We call them \textit{Patrons}, though they were never sworn to us. They are the tides that move unseen, the keepers of forgotten bargains, the sleepers beneath the stone and the stars. To call upon them is to dip a hand into a river that has carved mountains.''
\end{quote}

\subsubsection{The Patrons}
\index{Patrons}

Patrons are vast intelligences---not gods, though some worship them as such. They are embodiments of \textit{concepts} and \textit{forces} rather than sovereigns. Raéyn, mistress of the tides and the sea-routes. Khemesh, the crushing inevitability of the deep. Nidhoggr, the worm that dreams in the roots of time. Each offers power, but always with cost: fatigue, scars upon fate, or a slow unweaving of one’s own story.

To entreat a Patron is to risk being marked. Their Rites are gifts and snares both.

% Individual patrons (order however you like)
\input{patrons/carrion-king.tex}
\input{patrons/clockwork-monad.tex}
\input{patrons/gallows-bell.tex}
\input{patrons/grimmir.tex}
\input{patrons/ikasha.tex}
\input{patrons/inaea.tex}
\input{patrons/isoka.tex}
\input{patrons/khemesh.tex}
\input{patrons/livaea.tex}
\input{patrons/mab.tex}
\input{patrons/maelstraeus.tex}
\input{patrons/mykkiel.tex}
\input{patrons/nidhoggr.tex}
\input{patrons/oath-flame-light.tex}
\input{patrons/raeyn.tex}
\input{patrons/sacred-geometry.tex}
\input{patrons/sealed-gate.tex}
\input{patrons/traveler.tex}
\input{patrons/varnek-karn.tex}
\input{patrons/witness.tex}

\paragraph{No True Acolyte}
Interpreting a patron's will is often a dangerous prospect in and of itself. Many a Runkeeper has found themselves on the opposite end of machinations from others from the same patron.

\subsubsection{The Runekeepers}
\index{Runekeepers}

If Patrons are the storm, the \textbf{Runekeepers} are those who etched the first shelter. They do not serve; they remember. Their charge is to keep record of Rites, bindings, and the old words that tether meaning to symbol. A Runekeeper may never call a Rite themselves, but without their quiet stewardship, Invokers would stumble blind into bargains best forgotten.

\begin{quote}
``Every Rune is a promise. Every line a covenant. Do not mistake the Runekeeper’s silence for weakness; their memory is the foundation of our craft.''
\end{quote}

\subsubsection{The Invokers}
\index{Invokers}

\textbf{Invokers} are those who dare. Neither archivists nor worshippers, they are travelers on the knife-edge between story and ruin. An Invoker learns the Rites of a Patron, weaves them into their own Art, and bends fate for a moment. Yet invocation is not command: it is negotiation. The Patron always leaves its mark. The stronger the Rite, the deeper the scar.

Invokers are often wanderers, exiles, or seekers. To common folk they are feared---witches, oathbreakers, meddlers with things not meant for mortal hands. But when the village falls to plague, when the sea closes its roads, when the dead refuse their rest, it is an Invoker who is called upon.

\subsubsection{Closing Words}
The dance between Patron, Runekeeper, and Invoker is a triangle of peril and necessity. Without Patrons, there is no power. Without Runekeepers, no record. Without Invokers, no action. Together, they shape the crooked, perilous art we call Invocation.

\section{Patrons \& Rites}
\label{sec:patrons-rites}


% Optional: standard macro to format each Patron’s Gift call-out the same way
\newcommand{\PatronGift}[2]{% #1 = Thematic skill, #2 = brief domain blurb
\paragraph{Patron's Gift (Imbuement).}
Once per scene as an action (cost: 1 Boon; requires \textbf{Thiasos}), touch an item to imbue it until scene end with \textbf{+1 Melee} and \textbf{+1 #1}. \emph{Push It:} extend one more scene by marking \textbf{+1 Obligation}. Gifts from the same Patron don’t stack; take the best. Dice bonuses respect the +3 cap. \textit{Domain:} #2.
}

% Optional: standard macro to format each Patron’s Gift call-out the same way
\newcommand{\PatronGift}[2]{% #1 = Thematic skill, #2 = brief domain blurb
\paragraph{Patron's Gift (Imbuement).}
Once per scene as an action (cost: 1 Boon; requires \textbf{Thiasos}), touch an item to imbue it until scene end with \textbf{+1 Melee} and \textbf{+1 #1}. \emph{Push It:} extend one more scene by marking \textbf{+1 Obligation}. Gifts from the same Patron don’t stack; take the best. Dice bonuses respect the +3 cap. \textit{Domain:} #2.
}

\begin{table}[H]
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}p{3.8cm}p{3.8cm}p{7.5cm}@{}}
\toprule
\textbf{Patron} & \textbf{+1 Thematic Skill} & \textbf{Gift / Lore Bestowal} \\
\midrule
Ikasha (Shadow, Penumbra) & Stealth & Grants the hush between footsteps and the raven’s omen at every threshold. \\
Mykkiel (Judgment, Writ) & Command & Grants the authority of seal and sentence, words that bind like iron. \\
The Witness (Truth, Revelation) & Notice & Grants the unblinking gaze that unmasks deceit and remembers every oath. \\
Sealed Gate (Boundaries, Closure) & Tinker & Grants mastery of thresholds—doors that yield or bar at your command. \\
Raéyn (Storm, Tides) & Skirmish & Grants the sailor’s fortune: winds that shift, storms that answer to will. \\
Khemesh (Abyss, Pressure) & Skirmish & Grants the crushing silence of the deep, where strength is drowned in weight. \\
Mab (Glamour, Courts) & Persuade & Grants the mask of favor, a voice that bends courtiers and kindles desire. \\
Sacred Geometry (Perfect Forms) & Tinker & Grants the compass of perfection, every shape reduced to its true measure. \\
Clockwork Monad (Mechanism, Process) & Tinker & Grants the certainty of repetition: a cycle that never falters, a gear that never slips. \\
Varnek Karn (Ossuary, Dominion of the Dead) & Command & Grants the silence of the archive, where the dead obey and records speak. \\
Nidhoggr (Deep Earth, Rot) & Skirmish & Grants the weight of ages, the strength of stone and the hunger of roots. \\
The Traveler (Ways, Roads) & Notice & Grants the open way, a compass that never rests, and roads where none are marked. \\
Oath of Flame \& Light (Dawn, Vows) & Command & Grants the fire of dawn, a vow that shields the faithful and sears the faithless. \\
Carrion King (Carrion, Renewal) & Survival & Grants the feast of decay, where what is dead becomes seed for what lives. \\
Gallows Bell (Doom, Last Rites) & Command & Grants the toll of ending, a voice that closes stories and calls debts due. \\
Old Man of the Black Forest (Primal Humanity, Instinct) & Survival & Grants the wild memory: fang, fire, and the path of instinct through the dark wood. \\
Isoka (Serpents, Shedding) & Skirmish & Grants the serpent’s coil, strength in sudden strike and wisdom in renewal. \\
Inaea (Mercy, Hearth) & Persuade & Grants the hearth’s warmth, shelter to the weary and mercy for the lost. \\
Maelstraeus (Infernal Bargainer) & Persuade & Grants the contract’s weight, every deal sealed in fire and shadow. \\
Livaea (Temptation, Desire) & Persuade & Grants the lure of longing, beauty sharpened into power over hearts. \\
\bottomrule
\end{tabular}
\caption{Patron’s Gift: fixed Thematic Skill and lore of their bestowed blessing. Thematic bonuses apply only when the fiction matches the Patron’s domain.}
\label{tab:gift-thematic-map}
\end{table}

%========================================
% appendix/patron-rivalries-expanded.tex
%========================================
% Include with: \input{appendix/patron-rivalries-expanded.tex}
% Requires: \usepackage{booktabs,longtable}

\section*{Patron Rivalries}
\label{app:patron-rivalries}

Use this matrix to quickly shade rulings. “Edge Loci” are environments or situations where one side tends to start a step better in Position or gains an Effect nudge (Keeper’s call). “Friction” are handy prompts for SB spends.

\renewcommand{\arraystretch}{1.15}
\setlength{\LTpre}{0pt}
\setlength{\LTpost}{0pt}

\begin{longtable}{@{}p{3.3cm}p{3.3cm}p{4.6cm}p{7.2cm}@{}}
\toprule
\textbf{Patron} & \textbf{Rival} & \textbf{Edge Loci} & \textbf{Friction \& Prompts (SB)} \\
\midrule
\endfirsthead

\toprule
\textbf{Patron} & \textbf{Rival} & \textbf{Edge Loci} & \textbf{Friction \& Prompts (SB)} \\
\midrule
\endhead

\bottomrule
\endfoot
Raéyn (Sea, Tides, Travel) & Khemesh (Abyssal Maw) &
Open water, coasts, shipping lanes, storms you can \emph{read}. &
SB: changing tides; shifting winds; a clear route opens \emph{but} a vow at sea is invoked; waymarks appear then vanish; safe harbor demands a price. \\

Khemesh (Abyssal Maw) & Raéyn (Sea, Tides, Travel) &
Trenches, lightless holds, flooded caverns, oppressive silence. &
SB: pressure crush; voices from the bilge; hull-groan clocks; lanterns dim; maps untrustworthy; a crewman hears the trench call. \\

Sealed Gate (Boundaries, Closure) & The Traveler (Ways, Roads) &
Customs houses, oaths, locks, court thresholds. &
SB: writ checked; stamp demanded; wrong ledger; crossing exacts a toll; shortcut collapses into lawful detour; [WARD] keys hum. \\

The Traveler (Ways, Roads) & Sealed Gate (Boundaries, Closure) &
Desire paths, smuggler tracks, wayshrines, liminal crossings. &
SB: desire line opens; escort looks away; marginalia proves true; the lock refuses a lawful key \emph{now}. \\

The Witness (Truth, Revelation) & Mab (Glamour, Courts) &
Depositions, confessionals, cold light, mirrored chambers. &
SB: mask slips; testimony contradicts a courtier; illusion seams show; a polite scandal erupts. \\

Mab (Glamour, Courts) & The Witness (Truth, Revelation) &
Masques, salons, petty courts, festive oaths. &
SB: a favor called; duel by slight; truth offends protocol; a boon if the mask stays on. \\

Ikasha (Shadow, Latent Potential) & The Witness (Truth, Revelation) &
Deep shade, empty rooms, places holding unrealized action. &
SB: hush worsens Position under scrutiny; a shadow remembers your step; a deferred answer comes due. \\

Mykkiel (Judgment, Writ) & Varnek Karn (Necromantic Archives) &
Courts martial, audit halls, sanctified ledgers. &
SB: the writ binds the restless; precedent rejected; a ledger page missing; sentence invites a haunting. \\

Varnek Karn (Necromantic Archives) & Oath of Light \& Flame (Dawn, Vows) &
Ossuaries, plague pits, memorial crypts, last testaments. &
SB: bone answers but asks payment; unfinished business drags PCs into an old feud; consecration threatens the archive. \\

Oath of Light \& Flame (Dawn, Vows) & Khemesh (Abyssal Maw) &
Sunrise rites, consecrated decks, sworn escorts. &
SB: dawn burns back the hush; a vow compels aid; the abyss recoils \emph{but} leaves a later omen. \\

Ikasha (Shadow, Latent Potential) & Oath of Light \& Flame (Dawn, Vows) &
Twilight thresholds, watchfires, taboos kept at night. &
SB: light corners a secret; a vow drags the hidden into view; shadow offers escape at a cost. \\

Varnek Karn (Necromantic Archives) & Carrion King (Carrion, Renewal) &
Catacombs, bone-fields, funeral workshops. &
SB: ledger of skulls disputes the feast; remains refile themselves; scavengers desecrate protocol. \\

Maelstraeus (Infernal Bargainer) & Oath of Light \& Flame (Dawn, Vows) &
Witness trees, oath-stones, sanctified doors. &
SB: contract vs. vow; a pledge sears a loophole shut; a promised price comes due \emph{now}. \\

Livaea (Temptation, Desire) & Inaea (Mercy, Hearth) &
Bedchambers, feasting halls, clandestine trysts. &
SB: desire strains hospitality; temptation corrodes bonds; mercy offers comfort, but passion demands more. \\

The Traveler (Ways, Roads) & Clockwork Monad (Iteration, Process) &
Caravans, milestones, scheduled relays. &
SB: timetable buckles; a detour saves time; a routine yields speed \emph{but} blinds to ambush. \\

Nidhoggr (Dreaming Antiquity) & The Witness (Truth, Revelation) &
Ancient dreams, truth-seeing rites among ruins. &
SB: revelation wakes something old; truth contradicts myth—or proves it; an ancestor demands to be heard. \\
\bottomrule
\end{longtable}

\paragraph{Quick Rulings.}
\begin{itemize}
  \item \textbf{Position Nudge:} In a home locus, start one step better; in a rival locus, one step worse.
  \item \textbf{Effect Shade:} Where a Patron dominates, consider an Effect bump; where opposed, consider Limited Effect unless paid for.
  \item \textbf{Symbol Interference (Invokers):} Carrying both sides’ Symbols increases narrative noise: first ritual each scene may mark +1 Obligation (Keeper’s call).
  \item \textbf{SB Color:} When spending SB in these matchups, prefer suits that fit: Hearts (social), Spades (harm/escalation), Clubs (material cost), Diamonds (numinous disturbance).
\end{itemize}

% =========================
% OBLIGATION OVERFLOW (RITES)
% =========================
\section{Obligation Overflow (Rites)}
\label{sec:obligation-overflow}

When \textbf{Obligation} is ticked past its maximum (from Rites, vows, bargains), mark \textbf{Fatigue} based on scene severity:
\begin{itemize}
  \item \textbf{Low}: +1 Fatigue
  \item \textbf{Standard}: +2 Fatigue
  \item \textbf{High}: +3 Fatigue
\end{itemize}
If this \emph{fills} the Fatigue Track, apply the \textbf{Fatigue $\rightarrow$ Harm} conversion (see \S\ref{sec:health-fatigue-harm-rev}).
