\section{Isoka --- Angel of Serpents}
\label{patron:isoka}

\subsection*{Lore}
\index{Patrons!Isoka}%
Isoka, sister to Ikasha (Shadow) and Inaea (Mercy), completes the Triad of Transformation. Where thresholds and compassion mark her sisters' domains, Isoka teaches that every self is temporary---identity is a skin to be shed so growth can continue. Her serpents are omens and teachers: each cast skin a lesson in release; each venom a catalyst for necessary change. Those who walk her path become alchemists of the self, embracing dissolution as the first motion of rebirth.

\begin{quote}
``Do not mourn the skin you shed. It was never meant to last. The venom that burns away the old self is the same that grants the strength to become new.''
\end{quote}

\subsection*{Patron's Gift: Serpent's Grace}
Once per scene as an action (cost: 1 Boon; requires Thiasos), touch a weapon or tool to imbue it until scene end. The target gains +1 die and +1 Effect when used for transformation, deception, or striking from concealment.  
\textbf{Push It:} Extend for one additional scene by marking +1 Obligation. Attuned observers notice the serpent's favor.

\subsection*{Low Rites}
\paragraph{Rite of the Loosening Skin (Low)}%
\emph{Duration: Scene; Range: Self. Materials: Discarded snakeskin or loose thread.}  
Gain +1 die to resist an ongoing Condition this scene, or reroll one \textbf{1} on an escape/evasion. On success, you may declare the Condition \emph{shed} and create a 2-segment \emph{Transformation Residue} clock to ignore a similar effect later.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Also ignore one minor movement penalty; leave a token of the old self that can be traced (mark 1~SB \emph{Diamonds}).

\paragraph{Rite of the Subtle Shift (Low)}%
\emph{Duration: Scene; Range: Self. Materials: Palmed trinket passed hand-to-hand.}  
Fluid demeanor: +1 die to \emph{Deceive} to pass as a nearby class/profession, or +1 Effect to blend into a new crowd/site. Create a 4-segment \emph{Blended Identity} clock to downgrade one social complication.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Bypass one minor identity check; you must maintain the false role to scene end (generate 1~SB \emph{Hearts} if challenged).

\subsection*{Standard Rites}
\paragraph{Rite of the Shedding [TRANSFORM] (Standard)}%
\emph{Duration: Scene; Range: Self. Materials: Full change of clothing and an adopted mannerism.}  
+2 dice to resist one named ongoing Condition; once/session declare a minor physical contingency retroactively (``I packed the tool,'' ``I took that step earlier''). Create a 6-segment \emph{Shed Identity} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Clear a temporary identity-based Minor Condition; your former identity stirs in the fiction (ally, rival, or witness appears).

\paragraph{Rite of the Forked Tongue [BIND] (Standard)}%
\emph{Duration: Scene; Range: Self/Near. Materials: A harmless lie told to a mirror.}  
Ambiguous persuasion: when you \emph{Sway} or \emph{Command}, a success may generate \emph{Diamonds} (leverage) instead of SB. Targets of deception must test (Wits+Insight DV~3) or suffer --1 die to future interactions with you this scene. Create a 4-segment \emph{Verbal Venom} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} One carefully worded lie this scene is treated as true; the displaced truth seeks return (mark 1~SB \emph{Hearts}).

\subsection*{High Rites}
\paragraph{Rite of Complete Metamorphosis [TRANSFORM][WARD] (High)}%
\emph{Duration: Scene; Range: Self. Materials: Identity kit (garb, voice, tokens) and a serpent's shed skin.}  
Full appearance/voice change. Begin \emph{Controlled} on \emph{Deceive}/\emph{Stealth}; once/scene declare a minor contingency retroactively. You are [\emph{WARD}]ed against recognition by former acquaintances. Create an 8-segment \emph{New Identity} clock.  
\textbf{Invoke:} 1 action; mark +2 Obligation.  
\textbf{Push It:} Spoof scent/biometric once; your original identity partially unmoors and acts independently (mark 2~SB \emph{Spades/Hearts}).

\paragraph{Rite of the Cast-Off History [UNWARD][CURSE] (High)}%
\emph{Duration: Extended; Range: Self. Materials: Burning or defacing mundane records of the old life.}  
Upon completion, common records and casual memories of that identity become unreliable; trackers via that identity suffer --2 dice (magic and intimates still apply). Gain +2 dice to rolls with the new identity. Create a 6-segment \emph{Erased Past} clock.  
\textbf{Invoke:} Extended ritual; mark +2 Obligation.  
\textbf{Push It:} A plausible ``death'' occurs for the old identity; one intimate senses deception but cannot prove it (mark 1~SB \emph{Clubs}).

\subsection*{Obligation Progression}
Starts at 6 for Tier II characters, scaling with tier.

\paragraph{Obligation 9+} Isoka demands a significant transformation (abandon a cherished role, betray a former allegiance, or shed a defining trait). Refusal: GM immediately spends 2~SB to introduce identity complications.  

\paragraph{Obligation 11+} Your selfhood frays. Mark Permanent Condition \emph{Fragmented Self} (--2 dice to rolls requiring consistent personality or established relationships) until a quest of reintegration is completed.

\subsection*{Persistent Condition: Serpent's Fluidity}
Gain +2 dice to disguise, deception, escape, and identity transformation; suffer --1 die to rolls requiring consistent identity or long-term commitments. Narrative: adaptive and elusive, but hard to trust.

\subsection*{Rivalries}
\begin{itemize}
\item \textbf{The Pale Shepherd:} Tension---Isoka transforms the living; the Shepherd tends endings.  
\item \textbf{Maelstraeus:} Opposition---fluid selves vs.\ contractual fixation.  
\item \textbf{The Inquisitor Prime:} Antagonism---transformative ambiguity vs.\ purifying certainty.  
\end{itemize}

\subsection*{Serpent Mysticism}
Isoka’s three gifts: \emph{The Shedding} (release what no longer serves), \emph{The Venom} (destroy to renew), \emph{The Coil} (wait for the precise moment). Rites often employ serpent symbols; some devotees keep serpent familiars. The ``forked tongue'' is the art of holding multiple truths for different audiences.

\subsection*{Playtest Scenario: The Alchemist's Last Formula}
A master alchemist’s elixir grants any ability at the cost of permanently shedding an aspect of identity. Factions vie to claim it.  
\begin{itemize}
\item \emph{Loosening Skin} resists identity-loss shock.  
\item \emph{Subtle Shift} infiltrates rival groups.  
\item \emph{Complete Metamorphosis} enables the endgame persona swap.  
\item \emph{Forked Tongue} maneuvers factions with tailored truths.  
\item \emph{Cast-Off History} obscures the party’s involvement afterward.  
\end{itemize}
Resolution: destroy the formula, claim it and change forever, or engineer a merciful transformation that honors identity’s value while satisfying Isoka’s demand for becoming.
