# % --- Patron: Livaea, the Crimson Courtier (Seduction & Social Binding) ---

\subsubsection{Livaea, the Crimson Courtier (Seduction \& Social Binding)}
\textit{Lore.} In salons where wine flows like honey and words cut sharper than daggers, the Crimson Courtier holds court. She is the patron of those who would bind others not with webs, but with desire, obligation, and the sweet poison of whispered promises. Her followers are masters of the intimate covenant, the secret alliance, and the kiss that seals a fate. She teaches that the deepest wounds are those inflicted through trust, and the strongest chains are those forged from willing hands.

\begin{quote}
A word can wound deeper than a blade. A promise can chain stronger than iron. The Courtier knows which words to whisper—and which silences to sell.
\end{quote}

\paragraph*{Rite of the Velvet Whisper (Low, 4 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A silk handkerchief or ribbon touched to lips. \\
\textbf{Effect:} Your next whispered words carry supernatural weight; +1 die to Sway when speaking privately to one target. \\
\textbf{Invoke:} 1 action; mark +1 Obligation. \\
\textbf{Push It:} The target feels compelled to whisper back a secret of their own, but you mark Exposure +1 and the exchanged confidence creates \textbf{1 SB (Hearts)} as gossip spreads. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph*{Rite of the Intimate Covenant (Low, 5 XP)} \emph{Scene; Touch; No.}
\textbf{Materials:} A shared cup of wine or exchange of personal tokens. \\
\textbf{Effect:} Create a temporary bond of trust; both parties gain +1 die when cooperating, and suffer -1 die when acting against each other this scene. \\
\textbf{Invoke:} 1 action; mark +1 Obligation. \\
\textbf{Push It:} The bond becomes slightly magical - one party feels the other's emotional state, but you mark Fatigue 1 and the emotional intimacy leaves both parties vulnerable—mark \textbf{1 SB (Diamonds)} as psychic resonance lingers. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Binding Vow (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A ring or token held while speaking the vow. \\
\textbf{Effect:} Forge a magical agreement between willing parties; +1 Effect when working together, breach forces 1 SB (Hearts) on breaker. \\
\textbf{Push It:} The vow becomes supernaturally enforced - breaker suffers Harm 1 and cannot act against the agreement for one scene; the Court takes note of the binding—mark \textbf{1 SB (Clubs)} as social forces align around the vow. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Court's Favor (Standard, 7 XP)} \emph{Scene; Self; No.}
\textbf{Materials:} Perfumed oil or cosmetic applied before social interaction. \\
\textbf{Effect:} Gain +2 dice to social manipulation in refined settings; you appear perfectly attuned to the social environment. \\
\textbf{Push It:} Become the center of attention - all social actions in the scene focus on you, but you cannot leave unnoticed and attract unwanted admirers—mark \textbf{1 SB (Spades)} as social complications arise. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Crimson Alliance (High, 13 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A circle of red candles, each representing a participant. \\
\textbf{Effect:} Bind multiple parties in a web of mutual obligation; all participants gain +1 die when acting in group's interest, and suffer Harm 1 if they act against it. \\
\textbf{Push It:} The alliance becomes magically permanent - breaking it requires a ritual and advances a 6-segment "Broken Bonds" clock; the Court's attention intensifies—mark \textbf{2 SB (Hearts)} as social forces take notice. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\paragraph{Rite of the Eternal Court (High, 14 XP)} \emph{Extended; Zone; No.}
\textbf{Materials:} A throne or seat of honor consecrated with rare perfumes. \\
\textbf{Effect:} Establish yourself as the center of a social web; for the next session, all social interactions in your presence are influenced by your will (+1 Effect to your social actions, -1 die to those opposing you). \\
\textbf{Push It:} The court becomes supernaturally compelling - all who enter must test Resolve (DV 3) or become devoted to you for the scene; the Court's influence expands—mark \textbf{2 SB (Diamonds)} as supernatural social pressure affects the wider area. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments.

\subsection*{Livaea's Corruption Table}
\label{sec:livaea-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Silver Tongue: +1 die to Persuade or Sway in any social situation. & Compulsive Charm: Must make an effort to appear uncharming in casual interactions, or suffer -1 die to rolls requiring subtlety. \\
\midrule
2 & Heart's Insight: Once per scene, gain +2 dice to read someone's true intentions or hidden desires. & Emotional Vampire: Suffer 1 Fatigue when spending extended time away from social interaction or romantic attention. \\
\midrule
3 & Irresistible Presence: Allies within Near range gain +1 die to social rolls when you support them. & Attention Magnet: Cannot move through crowds unnoticed; always attract admirers, rivals, or curious onlookers. \\
\midrule
4 & Binding Words: Once per session, make a promise so compelling that the target must make a Resolve test (DV 4) or accept it unquestioningly. & Temptation's Burden: Suffer -1 die to resist诱惑s or offers that appeal to your vanity or desires. \\
\midrule
5 & Court's Favor: Once per session, call upon a network of social connections to gain information, favors, or assistance. & Social Dependency: Suffer 1 Fatigue when isolated or cut off from your social network for extended periods. \\
\midrule
6+ & Irresistible Ascendancy: Once per session, become the absolute center of attention in any social gathering. All present must make a Resolve test (DV 5) or become devoted to you for the scene. & Addiction to Adoration: Mark +2 Obligation when using this power; you become dependent on constant praise and attention, suffering severe penalties when deprived. \\
\bottomrule
\end{longtable}