% --- Patron: Mab, Queen of Courts (Glamour & Bargain) ---
\subsubsection{Mab, Queen of Courts (Glamour \& Bargain)}
\textit{Lore.} The blush of truth, the dagger of etiquette, the smile that writes debts in perfume. Mab rules where desire dresses itself as courtesy.

\begin{quote}
Bend, don’t bow. Smile, don’t promise.
\end{quote}

\paragraph*{Rite of the Trickster’s Bargain (Low)}
\textbf{Tags:} [BIND][SOCIAL] \\
\textbf{Duration:} Scene \\
\textbf{Range:} Near \\
\textbf{Materials:} A token freely given (ribbon, coin, flower, etc.) \\

\textbf{Effect:} Offer a bargain sealed by glamour. Target must choose: either accept the bargain (both gain +1 die on rolls aiding the deal’s fulfillment this scene), or refuse and immediately mark 1 Stress (Hearts). \\
\textbf{Invoke:} 1 action; mark +1 Obligation. \\
\textbf{Push It:} Both parties are magically bound — betrayal causes 1 Harm (Stress) and starts a “Bargain Broken” clock [4]. \\

\paragraph{Courtly Guise \textnormal{[VEIL]} (Low, 4 XP)} \emph{Action; Self; Yes (social only).}
\textbf{Materials:} Pin a sprig of green or silver thread.\\
\textbf{Effect:} Subtle glamour: \textbf{+1 die} to Persuade/Sway in refined settings; you appear as expected rank/guest.\\
\textbf{Push It:} Also mask one minor tell; the first piercing question in the scene generates \textbf{1 SB (Hearts)}.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Token of Favor (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A ribbon or ring bestowed.\\
\textbf{Effect:} Grant an ally \textbf{+1 die} to one social action against onlookers who recognize your favor; you gain \textbf{+1 effect} to support.\\
\textbf{Push It:} The token also chills a heckler (one beat of hesitation), but you mark \emph{Exposure +1}.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Mirror of Motives (Standard, 7 XP)} \emph{Action; Near; No.}
\textbf{Materials:} A polished shard or compact mirror.\\
\textbf{Effect:} Ask one pointed question about an NPC’s \emph{immediate} social goal; Keeper answers truthfully or with a strong tell. Gain \textbf{+1 die} to exploit it this scene.\\
\textbf{Push It:} Also expose a concealed slight or insult that matters to them, creating \textbf{1 SB (Hearts)} on that target.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{The Price Agreed \textnormal{[OATH]} (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Exchange a token of equal apparent value.\\
\textbf{Effect:} Bind a petty bargain (favor-for-favor). Breach forces \textbf{1 SB (Hearts or Diamonds)} on the breaker and stains their reputation locally this arc.\\
\textbf{Push It:} Sweeten terms with a minor boon (\(+1\) die once to the beneficiary), but you take \textbf{1 SB (Hearts)} if they later breach.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Sovereign Glamour \textnormal{[VEIL][REVEAL]} (High, 11 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A circle of green felt or silk.\\
\textbf{Effect:} Establish Court: allies in Zone gain \textbf{+1 die} to social actions; crude threats suffer \(-1\) die. Once, peel one disguise/illusion in Zone.\\
\textbf{Push It:} Name a \emph{Court Law} (e.g., no drawn steel): first violation \emph{forces 2 SB} on the violator.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 6 segments.

\subsection*{Mab's Corruption Table}
\label{sec:mab-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Glamour's Touch: +1 die to Deception or Performance when telling stories or lies. & Cannot knowingly speak a direct falsehood; must mislead through wordplay. \\
\midrule
2 & Fairy Step: Once per scene, instantly move Near range as if by teleport. & Cold Iron Allergy: Suffer 1 Fatigue when touching or struck by iron/steel. \\
\midrule
3 & Charm of Mischief: Spend 1 Boon to twist a Complication into a humorous or advantageous outcome. & Compulsive Trickster: Must play a prank or jest each session or mark 1 SB (Hearts). \\
\midrule
4 & Gift of Hospitality: Allies sharing food/drink you offer gain +1 die on Resolve rolls that scene. & Hospitality Bound: Cannot harm those who accept your bread/salt without marking +2 Obligation. \\
\midrule
5 & Fae Sight: You see invisible doors, veils, and illusions. Gain +2 dice to Notice supernatural glamours. & Truth Debt: When offered a “fair trade” you must either accept or mark 1 Fatigue resisting the urge. \\
\midrule
6+ & Crown of Twilight: Once per session, declare an Oath. For the scene, gain +2 dice to all rolls aligned with that Oath. & Oathbound: Breaking an Oath inflicts 1 Harm (Stress) and starts a GM-controlled “Oathbreaker” clock [6]. \\
\bottomrule
\end{longtable}