\section{Maelstraeus --- The Infernal Bargainer}
\label{patron:maelstraeus}

\subsection*{Lore}
\index{Patrons!Maelstraeus}%
Maelstraeus is the Merchant of Equities, the Infernal Bargainer who sits at the crossroads of every transaction in the cosmos. Neither wholly demon nor angel, but a conceptual force born from the first exchange---the moment when one thing was traded for another, and debt was born.  

He dwells in the pause between offer and acceptance. His realm is a vast marketplace where every deal ever struck echoes through eternity, where contracts signed in blood still hold power, and where the true price of everything is known---even if it is never spoken aloud.  

Maelstraeus embodies the principle that all value can be exchanged, that every relationship is transactional, and that the universe itself operates on a vast economy of favors, debts, and obligations. His followers learn that everything has a price, but also that everything can be bought.  

\begin{quote}
``All things have value. All values can be traded. The Merchant sees the true price---and always collects his due.''
\end{quote}

\subsection*{Patron's Gift: Merchant's Eye}
Once per scene as an action (cost: 1 Boon; requires Thiasos), you may touch an item or person to imbue it until the end of the scene. The target gains +1 die and +1 Effect when used in trading, negotiation, or value assessment.  

\textbf{Push It:} Extend for one additional scene by marking +1 Obligation. The cosmic ledger records the flow.

\subsection*{Low Rites}
\paragraph{Rite of the Fair Trade (Low)}  
\emph{Duration: Scene; Range: Near. Materials: A balance scale with equal weights.}  
Establish a neutral trading ground. All parties gain +1 die to negotiate in good faith. Create a 4-segment \emph{Equity Maintained} clock that can be spent to downgrade a social complication.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Compel one party to reveal their true terms; mark 1 SB (Hearts).

\paragraph{Rite of the Merchant's Eye (Low)}  
\emph{Duration: Scene; Range: Self. Materials: A foreign coin or token.}  
Gain +2 dice to appraise goods, judge value, or spot market opportunities. Create a 6-segment \emph{Market Insight} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Also sense emotional value of items, but mark Exposure +1 and 1 SB (Diamonds).

\subsection*{Standard Rites}
\paragraph{Rite of the Balanced Exchange [OATH] (Standard)}  
\emph{Duration: Scene; Range: Near. Materials: Two items of perceived equal value.}  
Facilitate a fair trade; both sides gain +1 Effect. If unfair, the disadvantaged party gains +2 dice to resist. Create a 6-segment \emph{Fair Dealing} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Enforce magically: breaking terms causes 1 SB (Hearts/Clubs).

\paragraph{Rite of the Contract Seal [BIND] (Standard)}  
\emph{Duration: Scene; Range: Touch. Materials: An official seal or stamp.}  
Mark an agreement with authority; +1 die to enforce, --1 die for those who break it. Contract becomes [BIND]ed---breaking it imposes --1 die to all social rolls for one session. Create an 8-segment \emph{Binding Agreement} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Breach ignites the document and inflicts Harm~1; mark 1 SB (Spades).

\subsection*{High Rites}
\paragraph{Rite of the Great Market [WARD][COMMAND] (High)}  
\emph{Duration: Scene; Range: Zone. Materials: A consecrated booth or trading post.}  
Create a zone of commerce. Allies gain +1 Effect on trades; may reroll one failed social roll. Enemies suffer --1 die to deception. Create a 10-segment \emph{Market Dominance} clock.  
\textbf{Invoke:} 1 action; mark +2 Obligation.  
\textbf{Push It:} Attract powerful allies and rivals; mark 2 SB (Hearts/Clubs).

\paragraph{Rite of the Cosmic Ledger [CLEANSE][CURSE] (High)}  
\emph{Duration: Extended; Range: Self. Materials: A book recording debts and credits.}  
Once per session, convert one resource into another (e.g.\ 1 Boon $\rightarrow$ 1 Fatigue). Settle debts at true cosmic value. Create a 6-segment \emph{Ledger Balance} clock.  
\textbf{Invoke:} Extended ritual; mark +3 Obligation.  
\textbf{Push It:} Make an imbalanced trade in your favor; create a 6-segment \emph{Karmic Debt} clock; mark 2 SB (Diamonds).

\subsection*{Obligation Progression}
Starts at 7 for Tier II characters, scaling with tier.

\paragraph{Obligation 10+} Maelstraeus appears to collect. Refusal marks you \emph{In Default of the Ledger}: --2 dice to trading/negotiation, 1 SB on economic rolls.  

\paragraph{Obligation 12+} Your name is written in the Great Ledger as debtor. Others pursue your debt, and transactions accrue cosmic ``interest.'' Requires a major quest to resolve.

\subsection*{Persistent Condition: Merchant's Sight}
Gain +2 dice on trading, negotiation, value assessment, and detecting deception in economic contexts. Suffer --1 die on genuine generosity or selfless acts. Narrative: You see relationships as transactions, but lose touch with true altruism.

\subsection*{Rivalries}
\begin{itemize}
\item \textbf{Morag the Hag:} Both bind desires through bargains---cunning versus cosmic economy.  
\item \textbf{The Witness:} Antagonistic---value exchange vs. truths that cannot be bought.  
\item \textbf{Livaea:} Friction---desire fulfilled through transaction versus temptation.  
\end{itemize}

\subsection*{Connection to Economic Principles}
Maelstraeus embodies the law of exchange: all value is relative and transactional. His rites reflect supply, demand, and hidden costs. The \emph{cosmic ledger} ensures debts are always balanced, creating tension between immediate gain and long-term accounting.

\subsection*{Playtest Scenario: The Merchant Prince's Gambit}
A merchant prince corners a vital resource market. The party must decide whether to break his monopoly or build a sustainable alternative.  

\begin{itemize}
\item Use \emph{Rite of the Fair Trade} to unite factions.  
\item Apply \emph{Rite of the Balanced Exchange} to secure fair agreements.  
\item Invoke \emph{Rite of the Great Market} to demonstrate alternatives.  
\item Expose debts with \emph{Rite of the Cosmic Ledger}.  
\item Seal new agreements with \emph{Rite of the Contract Seal}.  
\end{itemize}

Resolution: Choose to shatter the monopoly or reshape the economy---both with lasting cosmic consequences.
