\section{Mykkiel — Arbiter of the Writ}
\label{patron:mykkiel}

\subsection*{Lore}
\index{Patrons!Mykkiel}%
Mykkiel is the Arbiter of justice and keeper of sacred covenants, weighing speech against deed and sealing verdicts in cold iron. His sigil—balanced scales crossed by a sword—marks benches of judgment and sealed cells from the Sapphire Marches to the Sunward Courts. 

Mykkiel does not merely enforce law; he embodies the principle that justice requires both mercy and judgment. Law without compassion curdles into tyranny; compassion without structure dissolves into harm. His followers learn that true authority is forged from evidence weighed, precedent applied, and oaths honored.

\begin{quote}
``Name the charge. Name the terms. Then sign where you will bleed if you are wrong. The Word made manifest cannot be unsaid.''
\end{quote}

\subsection*{Patron's Gift: Arbiter's Authority}
Once per scene as an action (cost: 1 Boon; requires Thiasos), touch an item or person to imbue it until scene end. The target gains +1 die and +1 Effect when used in lawful proceedings, judgment, or authoritative command.\\
\textbf{Push It:} Extend for one additional scene by marking +1 Obligation. The court’s notice falls upon the scene.

\subsection*{Low Rites}

\paragraph{Rite of the Stamp of Authority (Low)}%
\emph{Duration: Scene; Range: Near. Materials: Cold-iron seal or writ-tag.}\\
Project visible legitimacy. Gain +1 die to Command/Sway when asserting a lawful claim or order; challengers suffer --1 die to resist. Create a 4-segment \emph{Legal Standing} clock that can be spent to downgrade a legal complication.\\
\textbf{Invoke:} 1 action; +1 Obligation.\\
\textbf{Push It:} A brief hush (one beat) stills hecklers; mark \emph{Exposure} +1 as higher authorities take interest.

\paragraph{Rite of Proper Notice (Low)}%
\emph{Duration: Scene; Range: Near. Materials: Writ-string tied and snapped.}\\
Name a lawful venue (dais, doorway, wagon). The first hostile act committed there suffers --1 die. Create a 6-segment \emph{Sacred Venue} clock protecting the designated area.\\
\textbf{Invoke:} 1 action; +1 Obligation.\\
\textbf{Push It:} Name a protected act (parley, surrender, testimony) gaining +1 Effect in the venue; breaking custom generates 1~SB (Hearts) and marks the breaker before the covenant courts.

\subsection*{Standard Rites}

\paragraph{Rite of the Writ of Compliance [COMMAND] (Standard)}%
\emph{Duration: Instant; Range: Near. Materials: Red cord knotted while speaking the order.}\\
Issue an immediate, simple command (``Stand down,'' ``Drop it,'' ``Open''). Target must comply or suffer a stated cost; DV by fiction (elites may test Resolve). On success, create a 4-segment \emph{Lawful Compliance} token to auto-succeed on a similar lawful command later.\\
\textbf{Invoke:} 1 action; +1 Obligation.\\
\textbf{Push It:} On compliance, impose --1 die on the target’s next aggressive act this scene; the order sets precedent—mark 1~SB (Spades).

\paragraph{Rite of the Speaking Seal [BIND] (Standard)}%
\emph{Duration: Scene; Range: Near. Materials: Wax seal impressed over a name or sigil.}\\
Sanctify a statement (truce, custody, claim). Contradictors suffer --1 die; you gain +1 die to enforce. If the seal is broken, the breaker suffers Harm~1 (Legal) and --2 dice to social rolls involving magistrates for one session. Create an 8-segment \emph{Binding Seal} clock.\\
\textbf{Invoke:} 1 action; +1 Obligation.\\
\textbf{Push It:} Once, ask who intends breach; the Keeper provides a strong clue or a name. Mark 1~SB (Diamonds) as truth draws covenant notice.

\subsection*{High Rites}

\paragraph{Rite of the Oath Irons [OATH][WARD] (High)}%
\emph{Duration: Extended; Range: Near. Materials: Two iron pins warmed, touched to wrists, then quenched.}\\
Bind two parties to a bounded term. Breach forces 2~SB and brands a faint iron-mark until amends. The oath is [BIND]ed to both; breaking it imposes --2 dice to all legal proceedings for one session. Create a 6-segment \emph{Sacred Oath} clock.\\
\textbf{Invoke:} Extended; +2 Obligation.\\
\textbf{Push It:} Extend to a small circle (up to four), each choosing one narrow exception (Keeper approves). Exploiting an exception generates 1~SB (Diamonds) as the covenant’s complexity invites scrutiny.

\paragraph{Rite of the Final Judgment [CLEANSE][CURSE] (High)}%
\emph{Duration: Extended; Range: Near. Materials: Complete record of a case, signed by recognized authority.}\\
Render a final, supernaturally enforced verdict. Target tests Spirit+Resolve (DV~5); on failure, full consequences apply without appeal. On success, limited appeal is possible, but the target suffers --2 dice on appeals for one arc. Create a 10-segment \emph{Divine Verdict} clock affecting all proceedings involving the target.\\
\textbf{Invoke:} Extended; +3 Obligation.\\
\textbf{Push It:} Make the verdict absolute and unappealable; create 2~SB (Hearts/Spades) as allies mobilize to overturn justice by other means.

\subsection*{Obligation Progression}
Starts at 6 for Tier II characters, scaling upward.

\paragraph{Obligation 9+} Mykkiel demands judgment where law and mercy clash or precedent fails. Refusal: your legal gifts falter for one session (--2 dice to lawful actions) and you generate 1~SB when asserting authority.

\paragraph{Obligation 11+} Law saturates your sight. Permanent Condition: \emph{Judge’s Eye} (--2 dice to informal or intimate interactions) until you complete a quest to \emph{learn when to look away}.

\subsection*{Persistent Condition: Arbiter’s Sight}
+2 dice to legal proceedings, judgment, and authoritative command; --1 die to acts of personal leniency or rule-bending. Narrative: you read the writ stamped upon every deed, yet the human cost grows heavy.

\subsection*{Rivalries}
\begin{itemize}
  \item \textbf{The Inquisitor Prime:} Tension—purity by ordeal vs.\ justice by process.
  \item \textbf{Morag the Hag:} Antagonism—cunning bargains vs.\ sanctioned oaths.
  \item \textbf{The Witness:} Opposition—revealed truths that unsettle precedent.
\end{itemize}

\subsection*{Covenant Courts and Magisterial Practice}
Mykkiel’s tradition rests on three pillars:
\begin{enumerate}
  \item \textbf{The Word:} Precise language of law that cannot be unsaid.
  \item \textbf{The Balance:} Evidence weighed and precedent applied.
  \item \textbf{The Seal:} Binding power that makes judgment manifest.
\end{enumerate}
Rites often involve writs, seals, oaths, and formal venues. Adherents serve as magistrates, advocates, clerks, and peace-brokers, maintaining archives of cases and precedent.

\subsection*{Playtest Scenario: The Merchant’s Trial}
A wealthy factor stands accused of ruining a rival by means lawful yet unjust. The quarter is split between letter and spirit of the law. The party must navigate factions and render or influence judgment.

\begin{itemize}
  \item \emph{Stamp of Authority} to establish standing in proceedings.
  \item \emph{Writ of Compliance} to still a tumultuous court.
  \item \emph{Speaking Seal} to bind testimony and terms.
  \item \emph{Oath Irons} to forge a settlement both parties must honor.
  \item \emph{Proper Notice} to sanctify the venue and prevent violence.
  \item \emph{Final Judgment} if a definitive precedent must be set.
\end{itemize}

\noindent Paths of resolution include strict legality, moral restitution, creative compromise, or procedural integrity—each testing Mykkiel’s balance of mercy and law.
