% --- Patron: Nidhoggr, the World-Worm (Dreaming Antiquity) ---

\subsubsection{Nidhoggr, the World-Worm (Dreaming Antiquity)}
\textit{Lore.} Beneath stone and sleep coils \textbf{Nidhoggr}, who gnaws at the roots of time. He does not speak quickly; he dreams in centuries. To press your ear to the earth is to risk drowning in the silence of aeons. Yet for those who endure, he whispers truths long buried, memories fossilized in stone, and the slow inevitability of cycles unbroken. His followers walk in twilight between dream and ruin, bearing the weight of all that has been.

\begin{quote}
Press your ear to the earth and wait. If it remembers you, it will answer.
\end{quote}

\paragraph*{Glimpse the Ancient’s Shadow (Low, 4 XP)} \emph{Action; Self; No.}\\
\textbf{Materials:} Dust ground from a weathered stone.\\
\textbf{Effect:} +1 die to interpret \emph{ancient} places, scripts, or artifacts; once this scene, ask one yes/no about the site’s original purpose.\\
\textbf{Push It:} Gain +1 Effect, but mark \emph{Fatigue 1} as stone’s patience weighs on you.\\
\emph{Requires: Familiar.}

\paragraph*{Drink from the Dreaming Deep (Low, 5 XP)} \emph{Instant; Self; No.}\\
\textbf{Materials:} Water poured over stone, swallowed with eyes closed.\\
\textbf{Effect:} Learn one hidden fact about the locale’s past; GM may reveal through echo or dream.\\
\textbf{Push It:} The vision lingers too clearly—gain an additional detail, but mark Exposure +1 and 1 SB (Clubs).\\
\emph{Requires: Familiar.}

\paragraph{Stone-Sleeper’s Murmur (Standard, 7 XP)} \emph{Scene; Near (touch locus); No.}\\
\textbf{Materials:} Ear pressed to bedrock, wall, or pillar.\\
\textbf{Effect:} Ask up to 3 questions about events once imprinted in this stone; answers are fragmentary but true.\\
\textbf{Push It:} One answer is delivered with precise sensory clarity; generate 1 SB (suit by Keeper).\\
\emph{Requires: Familiar + Codex.}

\paragraph{Awakened Chronicle (Standard, 9 XP)} \emph{Ritual; Zone; No.}\\
\textbf{Materials:} Chalk spiral and four local touchstones.\\
\textbf{Effect:} The zone replays a past moment in spectral echoes, visible to all. Witnesses gain +2 dice to Investigate/Recall about it.\\
\textbf{Push It:} Invoke a second memory from a different age; mark +1 Obligation.\\
\emph{Requires: Familiar + Codex.}

\paragraph{Dive into the World-Worm’s Dream (High, 12 XP)} \emph{Scene; Self; No.}\\
\textbf{Materials:} Circle of stones under open sky, lain upon in stillness.\\
\textbf{Effect:} Ask up to 3 questions about the \emph{distant past} or \emph{buried truths} here. Answers come as lucid dreams and omens.\\
\textbf{Push It:} The dream stretches into prophecy: gain +3 dice to one occult action, but mark 2 SB immediately.\\
\emph{Requires: Familiar + Codex + Tier III.}\\
\emph{Obligation:} 7 segments.

\paragraph{Eclipse of Aeons (High, 14 XP)} \emph{Extended; Zone; No.}\\
\textbf{Materials:} Stones from three ruins aligned in a circle.\\
\textbf{Effect:} Submerge a zone in deep-time resonance. For one session, history bleeds into present: ruins reform, shadows of the dead walk, and forgotten oaths stir. Allies gain +2 dice to Recall, Divination, or Investigation; enemies suffer -1 die when relying on the present alone.\\
\textbf{Push It:} The bleed becomes permanent until countered; mark +2 Obligation and start an 8-segment \emph{Time Fracture} clock.\\
\emph{Requires: Familiar + Codex + Tier III.}\\
\emph{Obligation:} 8 segments.

\subsection*{Nidhoggr’s Corruption Table}
\label{sec:nidhoggr-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Ancient Sight: +1 die to Lore when reading ruins or relics. & Temporal Drift: Speak or think in archaic patterns; -1 die in modern social dealings. \\
\midrule
2 & Stone Memory: Once/session, recall one precise fact of history as if witnessed. & Heavy Mind: Suffer 1 Fatigue when confronted with lies or distortions of history. \\
\midrule
3 & Earth’s Whisper: +2 dice to Notice when listening to stone or soil. & Root-Bound: -1 die to aerial or swift actions; feel tethered to ground. \\
\midrule
4 & Dreaming Insight: Once/scene, gain +1 die to actions tied to ancient mysteries. & Haunted Sleep: Dreams replay past ages; mark 1 SB (Clubs) if rest is denied. \\
\midrule
5 & World-Worm’s Gaze: Once/session, see through the earth to a distant ancient site. & Slow Pulse: -1 die to reactions requiring haste; act with ponderous inevitability. \\
\midrule
6+ & Aeons of Memory: Once/session, touch stone to access the full record of a place. Gain +3 dice to historical investigation. & Overload: Mark +2 Obligation and suffer 1 Harm (Stress); visions leave you vulnerable until they fade. \\
\bottomrule
\end{longtable}