# % --- Patron: Oath of Flame & Light (Dawn & Vows) ---

\subsubsection{Oath of Flame \& Light (Dawn \& Vows)}
\textit{Lore.} Light names, binds, and burns. The Oath favors sworn keepers—those who stand in the open and keep their word even when it costs.

\begin{quote}
Speak in the light. Keep it, or the light will keep \emph{you}.
\end{quote}

\paragraph*{Kindle Vow (Low, 4 XP)} \emph{Action; Self/Ally; Yes.}
\textbf{Materials:} Ampoule of consecrated spark.\\
\textbf{Effect:} Name a near-term pledge this scene (\emph{hold the line}, \emph{get them out}). Bearer gains \textbf{+1 die} to actions that keep it.\\
\textbf{Invoke:} 1 action; mark +1 Obligation.\\
\textbf{Push It:} First betrayal or hesitation \emph{forces 1 SB (Hearts)} on the bearer.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph*{Lay on Hands \textnormal{[CLEANSE][HEAL]} (Low, 5 XP)} \emph{Instant; Touch; No.}
\textbf{Materials:} Palm over wound; vow whispered.\\
\textbf{Effect:} Remove one minor affliction or downgrade \emph{Harm} by one step \emph{or} clear \emph{Fatigue 1}. DV by fiction for stubborn curses/poisons.\\
\textbf{Invoke:} 1 action; mark +1 Obligation.\\
\textbf{Push It:} Also grant \textbf{+1 die} to the target's next Resist this scene; you mark \emph{Exposure +1}.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Sunlit Parley (Standard, 7 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Vow-ring engraved with sunrise and true name.\\
\textbf{Effect:} Establish open terms: honest persuasion gains \textbf{+1 die}; deceit attempts suffer \(-1\) die in this scene's parley.\\
\textbf{Push It:} Once, demand a public answer; evasion \emph{forces 1 SB (Hearts)} on the evader.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Purge the Shadow \textnormal{[REVEAL][DISPEL]} (Standard, 9 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} Consecrated spark cracked to light.\\
\textbf{Effect:} Expose illusions/disguises and suppress one minor ongoing glamour/curse in Near (DV by fiction).\\
\textbf{Push It:} Also sear a lingering tell on the source; you can sense it once again this arc; mark \textbf{1 SB (Diamonds)}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Radiant Smite \textnormal{[FOLLOW-UP]} (Standard, 8 XP)} \emph{Action; Self; No.}
\textbf{Materials:} Consecrated spark smeared on weapon or badge.\\
\textbf{Effect:} Consecrate your next strike this scene. On your next successful \emph{melee} hit this scene:
\begin{itemize}
  \item Upgrade the hit's \textbf{Effect} by one step (to Great if applicable), and
  \item Add \textbf{+1 Harm (Burn)} \emph{or} force \textbf{1 SB (Spades)} on the target's side if the blow is narrative rather than wounding.
\end{itemize}
\textit{Special.} Versus Undead, Oath-breakers, or Outsiders: the blow also \emph{sears the untrue}. Undead/Oath-breakers suffer \(-1\) die on their next action; Outsiders gain \textbf{+1} segment on Leash/Exit Tally (Hit only). If the attack \emph{misses}, the smite charge lingers for one beat; after that it gutters, creating \textbf{1 SB (Diamonds)} as attention swells.\\
\textbf{Push It:} The strike flares—on hit, emit a Close burst: hostile creatures in Close suffer \(-1\) die for one beat \emph{or} are driven back (worse Position by one step). Mark \textbf{+1 Obligation}.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Covenant Blaze \textnormal{[OATH][FORTIFY]} (High, 12 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A brazier lit with three names spoken.\\
\textbf{Effect:} Those who swear within gain a halo: \textbf{+1 die} to acts that keep the oath; attackers against a haloed subject suffer \(-1\) die if the act would violate the sworn terms. Oath-breakers immediately \emph{force 2 SB (Hearts/Spades)} and the halo scorches them (1 Harm, Burn).\\
\textbf{Push It:} The blaze also sanctifies the threshold (one beat of temporary \texttt{[WARD]} vs.\ oath-breakers entering).\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.

\subsection*{Oath of Flame \& Light Corruption Table}
\label{sec:oath-flame-light-corruption}

\begin{longtable}{>{\raggedright\arraybackslash}p{1cm} p{5cm} p{5cm}}
\toprule
\textbf{Tier} & \textbf{Benefit} & \textbf{Cost / Quirk} \\
\midrule
1 & Oathbound Strength: +1 die to rolls when upholding a declared vow or protecting the innocent. & Rigid Honor: Must attempt to uphold vows even when tactically disadvantageous; suffer -1 die to rolls requiring flexibility. \\
\midrule
2 & Radiant Sight: Once per scene, gain +2 dice to detect deception, illusion, or supernatural corruption. & Blinding Truth: Suffer -1 die to social rolls involving subtlety or diplomatic deception. \\
\midrule
3 & Holy Flame: Gain +1 die to melee attacks against undead, demons, or oath-breakers. & Light's Burden: Suffer 1 Fatigue when forced to act in darkness or conceal your identity. \\
\midrule
4 & Unwavering Resolve: Once per session, treat a failed Resolve or Command roll as a success, but mark 1 SB (Hearts). & Moral Absolutism: Suffer -1 die to rolls involving moral ambiguity or compromise. \\
\midrule
5 & Dawn's Blessing: Once per session, heal all allies within Near range of Fatigue 1 and minor conditions. & Beacon of Light: Cannot hide your presence from supernatural detection; enemies seeking you gain +1 die. \\
\midrule
6+ & Avatar of Vows: Once per session, become the living embodiment of your oaths. For one scene, gain +2 dice to all rolls involving protection, truth, or justice, but breaking any vow causes immediate Harm 2 (Burn). & Sanctified Purpose: Mark +2 Obligation when using this power; your unwavering dedication makes you a target for those who would test your resolve. \\
\bottomrule
\end{longtable}