\section{Raéyn --- Mistress of the Sea}
\label{patron:raeyn}

\subsection*{Lore}
\index{Patrons!Raéyn}%
Raéyn is the tempestuous goddess of the sea, the restless tide that carries news between shores and the promise of change between lives. She is mother to all who sail, her voice the wind that fills sails and her moods the storms that test every mariner's resolve.

But Raéyn's heart is torn by her greatest tragedy: her son Khemesh, the Kraken of the Depths, who embodies the crushing inevitability of the ocean's dark heart. Where Raéyn brings change and opportunity, Khemesh brings the final, inescapable pressure that grinds all things to nothing. Sailors pray to Raéyn for safe passage and favorable winds, but whisper Khemesh's name when seeking to lay the dead to rest beneath the waves.

Raéyn is passionate, mercurial, and fiercely protective of those who respect her domain. She favors those who read currents, bargain with weather, and carry news between shores. But cross her, and the sea itself becomes your enemy: fair weather turns to fury, and every wave a judgment.

\begin{quote}
``Mark the tide, name your course, and trust the wave-road. But speak ill of Khemesh, and even I may let the deep take you.''
\end{quote}

\subsection*{Patron's Gift: Tide's Favor}
Once per scene as an action (cost: 1 Boon; requires Thiasos), you may touch a weapon, vessel, or item to imbue it until the end of the scene. The object gains +1 die and +1 Effect when used in maritime contexts or situations involving change, travel, or currents.  

\textbf{Push It:} Extend the blessing for one additional scene by marking +1 Obligation. The sea's attention becomes noticeable to other sailors.

\subsection*{Low Rites}
\paragraph{Rite of the Tidemark's Blessing (Low)}  
\emph{Duration: Scene; Range: Self. Materials: A knotted length of salt-twine brushed with seawater.}  
Treat slick, swaying, or water-slicked footing as stable for you this scene. Gain +1 die on boarding, balance, or shipboard movement. Create a 4-segment \emph{Tide's Favor} clock that can be spent to ignore one level of difficult terrain.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Extend to one ally in Close for one beat, but generate 1 SB (Spades: shifting deck/hazards).

\paragraph{Rite of the Whispering Currents (Low)}  
\emph{Duration: Instant; Range: Self. Materials: A shell held to the ear while facing the wind.}  
Learn the safest near-term route across water or coastline (reefs, eddies, patrols) or gain +1 die to navigation checks for this scene. If Khemesh's influence is present, suffer --1 die from conflicting currents.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Also learn the fastest route, but mark Exposure +1 (leaving a telltale wake).

\subsection*{Standard Rites}
\paragraph{Rite of the Changing Tide [PASSAGE] (Standard)}  
\emph{Duration: Scene; Range: Zone (water-adjacent). Materials: A handful of pebbles cast in a crescent.}  
Bias currents and water levels in the zone. Those moving with the tide gain +1 die; those moving against suffer --1 die. Small craft must test to hold position. Create a 6-segment \emph{Tidal Influence} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Brief surge or drawdown (one beat): open a ford or swamp a skiff; mark +1 Obligation.

\paragraph{Rite of the Wave-Road Blessing [WARD] (Standard)}  
\emph{Duration: Scene; Range: Route (sea-to-sea). Materials: Two sea-glass markers dropped overboard at start and end.}  
Consecrate a wave-road between two visible points. Allies gain +2 dice on travel, evade, or carry actions at sea. Designated pursuers suffer --1 die to intercept. One active wave-road at a time. Create an 8-segment \emph{Blessed Passage} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Extend the route's favor to an adjacent leg for one beat; mark +1 Obligation.

\subsection*{High Rites}
\paragraph{Rite of the Storm-Queen's Hand [AREA][FOLLOW-UP] (High)}  
\emph{Duration: Scene; Range: Zone (sea/shore/sky). Materials: A vial of rainwater gathered at three crossings.}  
Shape a storm-band over the zone. Choose two modes at cast; switch one once per scene:  
\begin{itemize}
\item \textbf{Propulsion:} Vessel gains +1 band of movement per beat (or +1 Effect to maneuvers).  
\item \textbf{Concealment:} Veil of rain/spray; ranged targeting impaired; --1 die to hostile sighting.  
\item \textbf{Smite:} Once per beat, lash with wave or lightning as [AREA] hazard.  
\end{itemize}
\textbf{Invoke:} 1 action; mark +2 Obligation.  
\textbf{Push It:} Add a third mode for one beat, then GM spends 1 SB on collateral; mark +1 Obligation.

\paragraph{Rite of the Mother's Wrath [BANISH][CURSE] (High)}  
\emph{Duration: Extended; Range: Zone. Materials: Tears of a betrayed lover mixed with salt from seven seas.}  
Curse those who wronged you. Target suffers --2 dice to maritime/weather rolls for one session. At sea, they must roll Spirit + Resolve (DV 4) each day or suffer Harm~1 (Weather). Create a 6-segment \emph{Mother's Ire} clock.  
\textbf{Invoke:} Extended ritual; mark +3 Obligation.  
\textbf{Push It:} Curse spreads to target’s allies/family; mark 2 SB (Diamonds).

\subsection*{Obligation Progression}
Starts at 6 for Tier II characters, scaling with tier.

\paragraph{Obligation 9+} Raéyn demands proof of devotion---navigate a dangerous passage, recover a lost treasure, or confront Khemesh's servants. Refusal causes all maritime rolls to suffer --2 dice and generate 1 SB when weather is involved.  

\paragraph{Obligation 11+} Khemesh notices you. You are hunted by his servants; deep water becomes perilous even under Raéyn's protection. Requires a quest to prove worth or appease both mother and son.

\subsection*{Persistent Condition: Child of the Tide}
Gain +2 dice on maritime travel, weather prediction, and navigation. Suffer --1 die on prolonged time away from the sea. The sea’s rhythm flows in your blood, making you exceptional at sea but restless on land.

\subsection*{Rivalries}
\begin{itemize}
\item \textbf{Khemesh:} Direct antagonism---mother’s change vs. son’s crushing pressure.  
\item \textbf{The Traveler:} Tension---fluid paths vs. fixed ways.  
\item \textbf{The Sealed Gate:} Opposition---Raéyn opens passages, Gates close them.  
\end{itemize}

\subsection*{Connection to Maritime Culture}
Raéyn’s rites emphasize the philosophy that the sea is not an obstacle but a partner. Her worship blends aid, hindrance, and the inevitability of change. The mother--son dynamic adds depth to coastal culture: Raéyn for the living, Khemesh for the dead.

\subsection*{Playtest Scenario: The Kraken's Gambit}
A trading fleet is trapped between pirates and Khemesh’s kraken-servants. The party must navigate the three-dimensional battlefield while appeasing Raéyn’s moods.

\begin{itemize}
\item Use \emph{Rite of the Changing Tide} to aid or hinder pursuit.  
\item Use \emph{Rite of the Wave-Road Blessing} to establish safe corridors.  
\item Invoke \emph{Rite of the Storm-Queen’s Hand} as a climactic storm.  
\item Curse a pirate captain with \emph{Rite of the Mother’s Wrath}.  
\end{itemize}

Resolution: The party must decide whether to appeal to Raéyn’s protection or broker peace between mother and son.
