\subsection{The Sacred Geometry}
\label{patron:sacred-geometry}

\paragraph{Lore.}
The Sacred Geometry is the mathematical expression of the universe's underlying structure --- the divine mathematics that governs all existence. It is the principle that reduces chaos to measure, that finds the golden ratio in petals and the perfect spiral in galaxies. Those who serve it understand that beneath apparent randomness lies immutable law, and that by mastering these patterns, one can bend reality to will.

Kon'reh is not merely a game but a sacred practice --- a way of seeing the world through the lens of perfect proportion. The board’s pieces represent fundamental forces, its movements echo cosmic harmonics, and mastery of its patterns grants insight into the universe's hidden order.

The Sacred Geometry does not create chaos but reveals it as illusion. Its followers are architects of certainty in an uncertain world, mathematicians of the divine who understand that every problem has a solution if one can only find the correct equation.

\emph{``Chalk, string, and a prayer to ratios. When the circle closes, luck remembers its place.''}

\paragraph{Patron's Gift (Imbuement): Divine Proportion.}
Once per scene as an action (cost: 1 Boon; requires \emph{Thiasos}), touch an object or space to align it with sacred ratios until scene end, granting \textbf{+1 die} to one related roll and \textbf{+1 Effect} when that action involves precision, balance, or mathematical principles.  
\emph{Push It:} Extend one extra scene by marking +1 Obligation. The alignment attracts obsessive attention from those sensitive to harmony.

% --- Low Rites ---
\subsubsection*{Low Rites}

\textbf{Rite of the Golden Mean (Low).}  
\emph{Duration:} Scene; \emph{Range:} Self; \emph{Materials:} A tool marked with the golden ratio ($\varphi \approx 1.618$).  
Gain +1 die to rolls requiring precision, balance, or proportion. On success, re-roll one die showing 1 or 2.  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Upgrade effect one step on a single roll; suffer -1 die to social rolls involving spontaneity for the scene.

\textbf{Rite of the Perfect Angle (Low).}  
\emph{Duration:} Scene; \emph{Range:} Touch; \emph{Materials:} Compass and straightedge consecrated in ritual.  
Treat difficult terrain, awkward positioning, or structural obstacles as one step easier this scene. Gain +1 die on spatial/architectural reasoning rolls.  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Extend benefit to one ally in Close range, but generate 1 SB (Clubs).

% --- Standard Rites ---
\subsubsection*{Standard Rites}

\textbf{Rite of the Harmonic Resonance [WARD] (Standard).}  
\emph{Duration:} Scene; \emph{Range:} Zone; \emph{Materials:} Geometric patterns drawn with precision.  
Create a zone of harmony. Outsiders crossing must test DV 3. On Hit: cross normally. On Partial: suffer -1 die inside. On Miss: cannot cross this beat.  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Fortify the pattern further but mark +1 Obligation.

\textbf{Rite of the Calculated Trajectory [REVEAL] (Standard).}  
\emph{Duration:} Scene; \emph{Range:} Self; \emph{Materials:} A perfect circle and a solved geometric problem.  
Gain +2 dice to prediction, trajectory, or pattern recognition. Ask two questions about mathematical relationships in the current scene.  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Predict one future event with certainty, but mark +1 Exposure.

% --- High Rites ---
\subsubsection*{High Rites}

\textbf{Rite of the Fundamental Equation [WARD][BIND] (High).}  
\emph{Duration:} Scene; \emph{Range:} Zone; \emph{Materials:} Complex diagram of universal constants.  
Declare one physics/magic rule different in the zone (no scene-ending absolutes; GM may veto). Once per scene, downgrade a Miss to Success \& Cost.  
\emph{Invoke:} 1 action; mark +2 Obligation.  
\emph{Push It:} Affect an adjacent zone for one beat; generate 2 SB.

\textbf{Rite of Kon'reh Mastery [OATH][FORTIFY] (High).}  
\emph{Duration:} Extended; \emph{Range:} Near; \emph{Materials:} A consecrated Kon'reh board and pieces representing fundamental forces.  
All participants make contested Wits + Lore. Winners gain +2 dice to strategy/pattern/logic rolls next session; losers suffer -1 die.  
\emph{Invoke:} Extended ritual; mark +3 Obligation.  
\emph{Push It:} Winner imposes one mathematical ``law'' for the session, but generate 2 SB (Diamonds).

% --- Traits ---
\paragraph{Obligation.} Starts at 5 (Tier II), scaling upward.

\paragraph{Hard Obligation Triggers.}  
\emph{Obligation 8+:} Demand to solve an impossible mathematical problem or suffer -2 dice to logic/pattern rolls until solved.  
\emph{Obligation 10+:} Connection to perfection erodes empathy: -2 dice to social rolls; Spirit + Resolve (DV 4) needed for chaotic/emotional situations.

\paragraph{Persistent Condition: Perfect Vision.}  
\emph{Effect:} +2 dice for pattern recognition, logic, or architecture; -1 die for creativity, improvisation, or emotional intuition.  
\emph{Narrative:} You see flawless solutions but struggle with humanity and flexibility.

\paragraph{Rivalries.}  
\emph{The Ninth Rim:} Opposition --- order vs. void.  
\emph{The Witness:} Tension --- truth revealed vs. certainty imposed.  
\emph{The Sealed Gate:} Friction --- defined boundaries vs. universal laws.

% --- Kon'reh ---
\paragraph{Connection to Kon'reh.}  
Kon'reh is the sacred game used to train perception of universal patterns. Each piece embodies a force:  
\begin{itemize}
  \item \textbf{Center:} The Monad (unity, origin)  
  \item \textbf{Orbitals:} Dyad through Decad (duality through completion)  
  \item \textbf{Movements:} Echo cosmic harmonics and progressions  
\end{itemize}
Advanced adepts use Kon'reh boards as ritual foci, with piece positions determining rite parameters.

% --- Playtest ---
\paragraph{Playtest Scenario: The Fractal Catastrophe.}  
A city district is plagued by ``fractal storms'' --- distortions of reality following recursive mathematical patterns. The party works with a Sacred Geometry adept to restore balance.  

Examples:  
\begin{itemize}
  \item \textbf{Golden Mean:} Navigates distorted space by finding stable proportions.  
  \item \textbf{Harmonic Resonance:} Establishes safe zones against distortions.  
  \item \textbf{Fundamental Equation:} Temporarily rewrites local physics.  
  \item \textbf{Kon'reh Mastery:} Climactic ritual game resolves the catastrophe.  
\end{itemize}
