\section{The Sealed Gate --- Keeper of Boundaries}
\label{patron:sealedgate}

\subsection*{Lore}
\index{Patrons!The Sealed Gate}%
The Sealed Gate is invoked when thresholds weaken, forbidden knowledge leaks, or Outsiders press against the walls of reality. It is the Abjurist’s Patron, embodying the principle that protection sometimes requires imprisonment, and safety may demand exile. Followers are warders and exorcists, but also philosophers of separation: those who believe the act of closing is sacred.

The Gate manifests as an armored figure, face hidden by a helm that shifts symbols: binding runes for warders, expulsion marks for exorcists, and an impassable refusal for transgressors.  

\begin{quote}
``You write borders into the world and prosecute trespass. Doors remember their true keepers; lines mean what you say they mean.''
\end{quote}

\subsection*{Patron's Gift: Gatekeeper's Authority}
Once per scene as an action (cost: 1 Boon; requires Thiasos), touch a weapon, tool, or barrier to imbue it until scene end. The target gains +1 die and +1 Effect when used to seal, bind, or prevent passage.  
\textbf{Push It:} Extend for one additional scene by marking +1 Obligation. The Gate’s presence is felt by would-be trespassers.

\subsection*{Low Rites}
\paragraph{Rite of the Sealed Threshold (Low)}%
\emph{Duration: Scene; Range: Touch. Materials: Chalk, wax, chain, or sigil.}  
Mark a threshold. Crossing parties suffer worsened Position or stumble on first entry (fiction decides). Create a 4-segment \emph{Boundary Maintained} clock to automatically fail one crossing attempt.  
\textbf{Invoke:} 1 action; +1 Obligation.  
\textbf{Push It:} Treat the threshold as difficult terrain; +1 Obligation cost to cross.

\paragraph{Rite of the Key’s Rebuke (Low)}%
\emph{Duration: Instant; Range: Near. Materials: Gesture or chain-clack.}  
Project a spectral hasp to stagger or disarm. On success, create a 2-segment \emph{Ward Active} token to auto-succeed on a similar defense later.  
\textbf{Invoke:} 1 action; +1 Obligation.  
\textbf{Push It:} Drop the object just beyond reach; +1 Obligation cost to retrieve.

\subsection*{Standard Rites}
\paragraph{Rite of the Circle of Denial [WARD] (Standard)}%
\emph{Duration: Scene; Range: Near. Materials: Salt, iron filings, or blessed chalk.}  
Outsiders crossing test DV = Cap. On Hit: cross but add DV to their Exit Tally; on Partial: +1; on Miss: fail this beat. Create a 6-segment \emph{Boundary Integrity} clock.  
\textbf{Invoke:} 1 action; +1 Obligation.  
\textbf{Push It:} Fortify circle; clearer tells; +1 Obligation.

\paragraph{Rite of the Writ of Passage [BIND] (Standard)}%
\emph{Duration: Scene; Range: Near. Materials: Spoken naming and scribed pass-mark.}  
Designate a route as permitted. Allies gain improved flow (Position/Effect bump). Unauthorized crossers suffer --1 die to movement. Create an 8-segment \emph{Authorized Passage} clock.  
\textbf{Invoke:} 1 action; +1 Obligation.  
\textbf{Push It:} Extend to extra ally/obstacle; +1 Obligation.

\subsection*{High Rites}
\paragraph{Rite of the Banishment Knot [BANISH][BIND] (High)}%
\emph{Duration: Instant; Range: Near. Materials: Knot sealed with gate-sigil.}  
Target an Outsider; test DV = Cap. On Hit: add DV to Exit Tally; on Partial: +1. If full, entity acts once then departs; cannot return for one session. Create a 4-segment \emph{Banishment Enforced} clock.  
\textbf{Invoke:} 1 action; +2 Obligation.  
\textbf{Push It:} Strip one tether/anchor or forbid threshold-crossing in this location for one session; +2 Obligation.

\paragraph{Rite of the Consecrated Barrier [WARD][UNWARD] (High)}%
\emph{Duration: Extended; Range: Zone. Materials: Relics from three faiths, iron bands, blood of a trespasser.}  
Consecrate area against unauthorized passage. Crossers test Spirit+Resolve (DV 4) or suffer Harm~1. Only proper authorization bypasses. Create a 10-segment \emph{Sacred Boundary} clock.  
\textbf{Invoke:} Extended; +3 Obligation.  
\textbf{Push It:} Make barrier permanent/fixed; start \emph{Boundary Maintenance} [6].

\subsection*{Obligation Progression}
Starts at 6 for Tier II characters, scaling upward.

\paragraph{Obligation 9+} The Gate demands you enforce a painful barrier (e.g., deny safe passage, seal a loved one away). Refusal: all sealing fails for one session, wards collapse, --2 dice to protection rolls.  

\paragraph{Obligation 11+} Boundaries haunt your perception. Permanent Condition: \emph{Prison of the Mind} (--2 dice to trust, intimacy, or crossing metaphorical lines) until you open yourself in quest.

\subsection*{Persistent Condition: Keeper’s Vigilance}
+2 dice to sealing, wards, and boundary enforcement; --1 die to trust, intimacy, or vulnerability. Narrative: unmatched protector, but difficult companion.

\subsection*{Rivalries}
\begin{itemize}
\item \textbf{The Traveler:} Antagonism---roads opened vs.\ thresholds closed.  
\item \textbf{The Ninth Rim:} Opposition---erosion of borders vs.\ their maintenance.  
\item \textbf{Morag the Hag:} Tension---both control passage; Gate by law, Morag by bargain.  
\end{itemize}

\subsection*{Boundary Philosophy}
Three pillars of the Gate’s teaching:  
\begin{enumerate}
\item \textbf{Marking:} Inscribe the limits of permission.  
\item \textbf{Enforcement:} Exert will to bar intrusion.  
\item \textbf{Consequence:} Understand seals create pressure that must be managed.  
\end{enumerate}

Followers specialize across physical, conceptual, temporal, or ontological boundaries. Exorcists invoke the Gate to return Outsiders to their realm.

\subsection*{Playtest Scenario: The Breached Sanctuary}
An ancient sanctuary built to contain a dangerous but not evil entity has been compromised by cultists. The party must decide whether to restore the seal, allow safe release, or destroy the being.  

\begin{itemize}
\item \emph{Sealed Threshold} to hold breaches temporarily.  
\item \emph{Circle of Denial} for party protection.  
\item \emph{Banishment Knot} to return entity.  
\item \emph{Consecrated Barrier} for lasting seal.  
\item \emph{Writ of Passage} to authorize approach.  
\item \emph{Key’s Rebuke} to defend against cultists.  
\end{itemize}

Resolution: re-seal (honoring the Gate), negotiate safe passage, or destroy the entity, each with unique complications and consequences.
