\section{The Traveler --- Guide of Ways}
\label{patron:traveler}

\subsection*{Lore}
\index{Patrons!The Traveler}%
The Traveler is the eternal guide of the road, guardian of those who walk the paths between what is and what might be. Among the Fhara caravans and Kuvani traders along the Way of Silk, the Traveler is invoked at every crossroads, honored with small offerings at each waypoint, and consulted before every major journey.  

The Traveler is not merely one who shows the way---they \emph{are} the way, existing in the pause between steps and in the choice of which path to take when roads fork. Every journey is both physical and spiritual; to move from one place to another is to transform, and the road itself becomes a teacher.  

\begin{quote}
``One foot in a promise, and the road will meet you halfway. But break your word to the way, and the way will break you.''
\end{quote}

\subsection*{Patron's Gift: Road's Blessing}
Once per scene as an action (cost: 1 Boon; requires Thiasos), touch an item or person to imbue it until the end of the scene. The target gains +1 die and +1 Effect when used for movement, navigation, or pathfinding.  

\textbf{Push It:} Extend for one additional scene by marking +1 Obligation. The road’s attention is noticed by other travelers.

\subsection*{Low Rites}
\paragraph{Rite of Road-Sense (Low)}  
\emph{Duration: Scene; Range: Self. Materials: Road-nail or waystone pebble.}  
Unerringly pick the fastest safe route in Near/Far. Gain +1 die to avoid ambushes or delays. Create a 4-segment \emph{Path Memory} clock to ignore difficult terrain once.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Spot one hidden bypass, but generate 1 SB (Clubs).

\paragraph{Rite of the Traveler's Boon (Low)}  
\emph{Duration: Scene; Range: Self/Ally. Materials: Thread tied at the wrist.}  
Ignore one level of difficult terrain or bureaucracy; +1 Effect to travel or escape checks. If shared, create a 2-segment \emph{Shared Journey} bond with an ally.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Extend to one more ally; mark 1 SB (Diamonds).

\subsection*{Standard Rites}
\paragraph{Rite of the Waymark [PASSAGE] (Standard)}  
\emph{Duration: Scene; Range: Near. Materials: Chalk mark or small cairn.}  
Declare a lane as permitted/easy. Allies gain improved Position/Effect or ignore one obstacle. Create a 6-segment \emph{Marked Path} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Lane persists between scenes; first enemy to exploit it forces 1 SB (Spades).

\paragraph{Rite of the Bridge Between [TRANSPORT] (Standard)}  
\emph{Duration: Instant; Range: Near. Materials: Two pinches of road-dust clapped.}  
Relocate a willing target within Far along a visible/named route. Unwilling targets may resist (Body+Resolve DV~3). Create a 4-segment \emph{Pathway Established} clock.  
\textbf{Invoke:} 1 action; mark +1 Obligation.  
\textbf{Push It:} Carry one extra ally or bundle; arrivals off-balance.

\subsection*{High Rites}
\paragraph{Rite of the Crown of Crossings [WARD][COMMAND] (High)}  
\emph{Duration: Scene; Range: Zone. Materials: Brass compass missing its needle.}  
Call the Road: allies gain +1 die to move/evade; pursuers suffer --1 die. Once, declare ``the long way is short'' to finish a travel clock segment for free. Enemies must test (Wits+Command DV~4) or suffer --1 die to movement.  
\textbf{Invoke:} 1 action; mark +2 Obligation.  
\textbf{Push It:} Seal a hostile route briefly; generate 2 SB (Clubs/Diamonds).

\paragraph{Rite of the Wayfarer's Covenant [OATH][FORTIFY] (High)}  
\emph{Duration: Extended; Range: Near. Materials: Waystones from multiple regions.}  
Bind present parties to safe passage. While honored: +2 Effect on travel, reroll one failed travel roll per scene. Breaking oath inflicts Harm~1 (Fatigue) and marks breaker as \emph{Oathbreaker of the Road} (–2 dice to travel rolls).  
\textbf{Invoke:} Extended ritual; mark +3 Obligation.  
\textbf{Push It:} Breaking the covenant inflicts Harm~2 (Stress) and attracts hostile Wayward spirits.

\subsection*{Obligation Progression}
Starts at 5 for Tier II characters, scaling with tier.

\paragraph{Obligation 8+} The Traveler demands you guide someone/something to a destination. Refusal: all routes fail, terrain counts as difficult, –2 dice to navigation.  

\paragraph{Obligation 10+} The road claims you. You cannot remain still: –2 dice in stationary contexts, Body+Resolve (DV~3) required to endure cities or confinement.

\subsection*{Persistent Condition: Child of the Road}
Gain +2 dice to movement, navigation, and pathfinding. Suffer –1 die to rolls requiring prolonged settlement or stillness. Narrative: restless blood, exceptional in travel, uneasy when stationary.

\subsection*{Rivalries}
\begin{itemize}
\item \textbf{The Sealed Gate:} Antagonism---the Traveler opens paths, the Gate closes them.  
\item \textbf{The Pale Shepherd:} Subtle tension---ways between places vs.\ ways between states.  
\item \textbf{The Ninth Rim:} Opposition---the Traveler connects, the Rim erases.  
\end{itemize}

\subsection*{Connection to the Way of Silk}
The Traveler is most honored along the Way of Silk:
\begin{itemize}
\item Waystone offerings at crossroads  
\item Path blessing rites before departure  
\item Road-sharing pacts between caravans  
\item Wayfarer's covenants ensuring aid on dangerous routes  
\end{itemize}
Symbols: the \emph{Crown of Crossings} (a traveler's hood/hat) and the compass without needle (truth that the path is not always obvious).

\subsection*{Playtest Scenario: The Caravan's Last Journey}
A great trading caravan attempts its final crossing of the Way of Silk. Bandits, storms, and ancient curses threaten its survival. The party must protect it.  

\begin{itemize}
\item Use \emph{Rite of Road-Sense} to bypass ambushes.  
\item Use \emph{Rite of the Waymark} to secure campsites and routes.  
\item Invoke \emph{Rite of the Crown of Crossings} against a supernatural storm.  
\item Perform \emph{Rite of the Wayfarer's Covenant} to unify caravan factions.  
\item Employ \emph{Rite of the Bridge Between} to relocate isolated supplies.  
\end{itemize}

Resolution: deliver the caravan to safety or take a profane shortcut---incurring the Traveler’s wrath.
