\subsection{Varnek Karn, the Bone King}
\label{patron:varnek-karn}

\paragraph{Lore.}
Varnek Karn is the Archivist of the Dead, Keeper of Unfinished Business, and Lord of Necrotic Ledgers. Beneath stone and sleep lies the vast archive where every dying breath becomes a debt, every unspoken word a contract, and every unfinished task a binding obligation. Bones remember not only who died, but what they owed, promised, and wished undone. Varnek maintains the cosmic ledgers that track these accounts across lifetimes, ensuring that nothing owed to the dead is forgotten—and nothing owed by the dead goes unpaid. Many necromancers have sought dominion over the dead, only to find themselves entered into Varnek’s ledgers as payment for hubris. The Bone King does not merely raise corpses; he enforces the economy of death.  
\emph{``Bones remember what souls forget. Every debt paid creates another debt owed.''}

\paragraph{Patron's Gift (Imbuement): Ledger's Authority.}
Once per scene as an action (cost: 1 Boon; requires \emph{Thiasos}), touch an item or person to imbue it until scene end with \textbf{+1 die} and \textbf{+1 Effect} when leveraging death-bound obligations, unfinished business, or necrotic debts.  
\emph{Push It:} Extend one extra scene by marking +1 Obligation. The ledger takes note of your manipulation.

% --- Low Rites ---
\subsubsection*{Low Rites}

\textbf{Rite of the Unfinished Ledger (Low).}  
\emph{Duration:} Scene; \emph{Range:} Touch; \emph{Materials:} A binding thread tied to remains or a personal relic.  
Learn one piece of unfinished business binding a spirit (name a task, debt, or oath). Acting on it grants +1 die once this scene to relevant rolls. Create a 4-segment \emph{Debt Tracked} clock.  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Also learn one adversary tied to that business; generate 1 SB (Hearts or Diamonds) as the ledger cross-references your involvement.

\textbf{Rite of the Whispered Debt (Low).}  
\emph{Duration:} Instant; \emph{Range:} Self; \emph{Materials:} A pinch of grave-dust stirred into breath.  
Gain +1 die to investigate a recent death (within a day) or ask one yes/no about the cause of death and any debts incurred. If the deceased owed a debt, you may temporarily assume it (mark 1 segment on a \emph{Temporary Debt} clock [6]).  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Learn one fleeting sensory shard (sound, scent, image) about the debt’s nature; mark +1 Exposure.

% --- Standard Rites ---
\subsubsection*{Standard Rites}

\textbf{Rite of Speaking Bones [REVEAL][BIND] (Standard).}  
\emph{Duration:} Scene; \emph{Range:} Touch; \emph{Materials:} Oil the jaw; set the teeth with a sigil.  
A corpse answers 2 questions about the circumstances of death from its own perspective (fragmented, literal). Max 1 corpse/scene. If it had unfinished business, you may bind it to your service (Cap 1–2), then mark 1 segment on a \emph{Ledger Attention} clock [6].  
\emph{Invoke:} 1 action; mark +1 Obligation.  
\emph{Push It:} Ask a third question; the spirit’s memories fracture—mark 1 SB (Diamonds) as necromantic resonance disturbs the area.

\textbf{Rite of the Debt Assumption [OATH] (Standard).}  
\emph{Duration:} Extended; \emph{Range:} Self; \emph{Materials:} Sign the ledger with blood while naming the debt to be assumed.  
Assume responsibility for another’s unfinished business or debt. Gain +2 dice to actions fulfilling that debt, but mark 2 segments on your \emph{Personal Debt} clock [8]. If you fail to fulfill within one arc, suffer Harm 2 (Curse) and -2 dice to all necromantic actions.  
\emph{Invoke:} Extended ritual; mark +2 Obligation.  
\emph{Push It:} Assume a greater debt for greater power; mark 3 segments on \emph{Personal Debt} and start a \emph{Ledger Scrutiny} clock [4].

% --- High Rites ---
\subsubsection*{High Rites}

\textbf{Rite of the Court of Echoes [WARD][COMMAND] (High).}  
\emph{Duration:} Scene; \emph{Range:} Zone; \emph{Materials:} Circle of bone-totems and a tolling chime.  
Convene an advisory court of the dead (treat as a Major Asset clock [6]) that can: answer up to 3 complex questions, sway the fearful, or impose silence upon casual lies in the zone. All present test Spirit + Resolve (DV 3) or suffer -1 die to deception for the scene.  
\emph{Invoke:} 1 action; mark +2 Obligation.  
\emph{Push It:} Command a decisive action (e.g., rout a mob, burst a door), then tick the court twice; mark 2 SB (Hearts) as spectral tensions rise.

\textbf{Rite of the Final Settlement [CLEANSE][CURSE] (High).}  
\emph{Duration:} Extended; \emph{Range:} Touch; \emph{Materials:} A complete accounting written in blood on bone.  
Settle all unfinished business for one deceased individual. Prevents undead reanimation and ensures peaceful rest. You must pay the settlement cost—assume the debts, offer equivalent service to Varnek, or sacrifice something of equal value. Start a \emph{Settlement Complete} clock [8] that protects the area from necromantic intrusion.  
\emph{Invoke:} Extended ritual; mark +3 Obligation.  
\emph{Push It:} Settle multiple individuals at once; the cost multiplies. Mark 2 SB (Spades) as the ledger audits your own account.

% --- Traits ---
\paragraph{Obligation.} Starts at 7 (Tier II), scaling upward.

\paragraph{Hard Obligation Triggers.}  
\emph{Obligation 10+:} Varnek appears to collect: fulfill a significant obligation, assume another’s major debt, or sacrifice equivalent value. Refusal: marked \emph{In Default} (−2 dice to necromancy; haunted by collectors).  
\emph{Obligation 12+:} Your name enters the Great Ledger as debtor to the Bone King. Necromancers seek to claim your debt; undead are hostile. Only a major quest can fulfill an impossible service or win forgiveness.

\paragraph{Persistent Condition: Ledger-Bound.}  
\emph{Effect:} +2 dice for necromancy, debt negotiation, unfinished business; −1 die to living social interactions.  
\emph{Narrative:} The dead recognize your authority; the living sense something “off” and distrust the keeper of death’s accounts.

\paragraph{Rivalries.}  
\emph{The Pale Shepherd:} Opposition — Varnek sustains the dead’s business; the Shepherd guides to rest.  
\emph{The Inquisitor Prime:} Antagonism — preserving necromantic obligations vs. purifying destruction.  
\emph{Morag the Hag:} Tension — obligations by cosmic ledger vs. obligations by cunning bargain.

\paragraph{Connection to Necrotic Practices.}
Varnek’s power grows with the complexity of the debt economy. Raising the dead without settlement adds to the ledger’s leverage. Wise practitioners work \emph{with} the ledger, assuming small debts for greater power while carefully managing obligations. Unfinished Business is central—every lingering spirit is bound by an unmet obligation. Understanding and settling these obligations enables authority over the dead without crude coercion.

\paragraph{Playtest Scenario: The Merchant’s Last Deal.}
A wealthy merchant died before completing a transaction that would have prevented war. His spirit remains restless while factions vie to claim his “debt.” The party must navigate obligations without becoming overindebted.  
\begin{itemize}
  \item \textbf{Unfinished Ledger:} Reveal the final deal and its binding terms.  
  \item \textbf{Debt Assumption:} Temporarily take on the obligation to negotiate peace.  
  \item \textbf{Speaking Bones:} Question the merchant for terms and the heirloom’s location.  
  \item \textbf{Court of Echoes:} Convene witnesses for the final settlement.  
  \item \textbf{Final Settlement:} Complete the deal and square the cosmic account with Varnek Karn.  
\end{itemize}
