% --- Patron: Vorthak, the Withered Crown (Curse & Blight) ---
\subsubsection{Vorthak, the Withered Crown (Curse \& Blight)}
\textit{Lore.} Vorthak bears the crown of thorns, the scepter of withered grain. He is the Patron of Unnatural Endings, the one who teaches that some gifts come at a price that festers. His followers learn to weave curses that grow stronger over time, to blight what others hold dear, and to find power in decay that feeds upon itself.

\begin{quote}
``Every blessing carries its shadow. Every joy, its thorn. I am the whisper that turns fortune to rot.'' — Vorthak, the Withered Crown
\end{quote}

\paragraph{Thorn's Gift (Low, 4 XP)} \emph{Scene; Touch; Yes (curse only).}
\textbf{Materials:} A wilted flower or rusted nail.\\
\textbf{Effect:} Bestow a minor curse that worsens over time. Target suffers \(-1\) die on one chosen action for this scene; each time they fail that action, the penalty increases by \(-1\) (max \(-3\)).\\
\textbf{Push It:} The curse spreads to a related action; mark \textbf{1 SB (Hearts)} as the blight seeks new targets.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Rite of the Withered Bloom (Low, 5 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A healthy plant twisted into a knot.\\
\textbf{Effect:} Cause one living thing to sicken and wither; plants/fungi lose vitality, creatures suffer Fatigue 1, crafted items become [COMPROMISED].\\
\textbf{Push It:} The blight spreads to similar targets nearby; mark \textbf{1 SB (Clubs)} as the decay accelerates unnaturally.\\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).}

\paragraph{Mark of the Thrice-Turned Spindle (Standard, 8 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} Thread wound counterclockwise around a broken spindle.\\
\textbf{Effect:} Curse a crafted item, location, or ongoing project. It suffers persistent bad luck: \(-1\) Effect on all related rolls, and complications arise more frequently (GM gains +1 SB budget for related actions).\\
\textbf{Push It:} The curse becomes contagious to similar works/projects; mark \textbf{1 SB (Diamonds)} as the blight gains malevolent intelligence.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{The Hungry Rot (Standard, 9 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} Rotten fruit buried in sacred earth.\\
\textbf{Effect:} Create a zone of supernatural decay. All healing/restoration actions suffer \(-1\) die; [HEALED] and [MENDED] tags cost double Boons; food/water spoils rapidly.\\
\textbf{Push It:} The rot spreads one zone outward and begins consuming structural integrity; mark \textbf{2 SB (Spades)} as the blight attracts other malevolent forces.\\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).}

\paragraph{Crown of Thorns \textnormal{[BANE]} (High, 12 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A circlet of twisted black iron bearing three cursed gems.\\
\textbf{Effect:} Bind a powerful curse to one target. They suffer \(-2\) dice to all actions for the scene and mark Fatigue 1 each time they attempt to resist the curse's effects.\\
\textbf{Push It:} The curse becomes hereditary/bonded to their bloodline or closest allies; mark \textbf{2 SB (Hearts)} as the blight takes root in their destiny.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 7 segments.

\paragraph{The Feast of Worms (High, 14 XP)} \emph{Extended; Region; No.}
\textbf{Materials:} A crown of bone set with the teeth of oath-breakers.\\
\textbf{Effect:} Curse an entire region with persistent misfortune. All rolls suffer \(-1\) die; clocks advance more quickly; 1 SB generated each session that cannot be banked (must be spent immediately on complications).\\
\textbf{Push It:} The curse evolves to target the caster's greatest strength, turning it to weakness; mark \textbf{3 SB (Diamonds)} as the blight achieves terrible sentience.\\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).}\\
\emph{Obligation:} 8 segments.

