% patrons/the-witness.tex
% Fate’s Edge — Patron: The Witness (Truth & Omen)

\subsubsection{The Witness (Truth & Revelation)}

\paragraph{Lore.}
The Witness remembers what others bury. Every shadow cast and oath broken is a line in her unending ledger. She is the keeper of inconvenient truths, the patron of those who seek to expose lies or recover forgotten knowledge. Her followers learn that knowledge comes with a price—the weight of remembering what others would forget.

\paragraph{Quote.}
\emph{``I will show you what you would rather forget. But first, you must forget what you think you know.''} — The Witness

% --- Revised Rites Below ---

\paragraph{Rite of the Lingering Glimpse (Low, 4 XP)} \emph{Instant; Near; Yes (Investigation/Notice only).}
\textbf{Materials:} A trace of the thing to be remembered (hair, dust, a spoken name). \\
\textbf{Effect:} Gain +1 die to your roll to investigate or notice something directly related to the trace within the current scene. \\
\textbf{Push It:} Gain +2 dice instead, but mark 1 segment on a \textbf{Memory Strain Clock [4]}. If the clock fills, you gain Fatigue 1 and suffer -1 die on Investigation/Notice rolls until the end of the next scene due to mental exhaustion from forced recall. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).} \\
\emph{Note:} The Witness grants fleeting, focused insight drawn from the faintest echoes. % Replaces Mark of Remembrance

\paragraph{Rite of Piercing Scrutiny (Low, 5 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A circle drawn with chalk or string while focusing on the truth to be sought. \\
\textbf{Effect:} Within the zone, gain +1 die to rolls to detect deception (Insight vs. Deceit, spotting social tells) or to recall hidden knowledge (Lore/Investigate for memory). Social interactions within the zone begin one Position step worse for those attempting to deceive. \\
\textbf{Push It:} One target within the zone must make a Wits test (DV 3) or involuntarily reveal one pertinent lie or hidden fact they are currently concealing (Keeper determines relevance). Regardless of the test result, mark Exposure +1 for the target(s) in the zone. \\
\emph{Requires: Familiar \ (\textit{Invoke:} 1 Boon).} \\
\emph{Note:} The Witness's gaze makes falsehoods brittle and memories more accessible. % Replaces Rite of Testimony

\paragraph{Rite of the Echoing Truth [OMEN] (Standard, 8 XP)} \emph{Instant; Near; No.}
\textbf{Materials:} A reflective surface (mirror, still water, polished metal) used to focus on the target. \\
\textbf{Effect:} Target must make a Resolve test (DV 3) or suffer -1 die to rolls involving memory, deception, or resisting interrogation for the scene. If they fail, you may ask one specific, factual question about something they know, and they must answer truthfully or suffer 1 SB (Hearts) as the memory is forcibly drawn forth. \\
\textbf{Push It:} If the target fails their Resolve test, you may ask a second question, but the mental intrusion causes them Harm 1 (Stress/Mental). \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).} \\
\emph{Note:} The Witness drags reluctant truths into the light. % Replaces Omen of Recall

\paragraph{Rite of the Immutable Record [OATH] (Standard, 7 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A document signed by all parties within the zone, or a spoken pact witnessed by the caster. \\
\textbf{Effect:} Bind the agreement. Any party who knowingly breaches it suffers 1 SB (Hearts) immediately and gains a persistent \textbf{Oathbreaker's Mark} Condition (-1 die on social rolls involving honor, trust, or oaths until amends are made or a significant act redeems them). \\
\textbf{Push It:} The bond becomes magically enforced for one specific, crucial clause: violation automatically inflicts Harm 1 (Stress) on the breaker in addition to the SB and Mark. \\
\emph{Requires: Familiar + Codex \ (\textit{Invoke:} 1 Boon).} \\
\emph{Note:} The Witness ensures promises carry weight. % Replaces The Written Ledger

\paragraph{Rite of the Unveiled Heart [OMEN] (High, 12 XP)} \emph{Scene; Near; No.}
\textbf{Materials:} A private setting where the target feels safe or is speaking freely. \\
\textbf{Effect:} The target suffers -2 dice to all attempts to conceal true emotions, intentions, or lies for the scene. Any successful social roll (Sway, Command, Deceit) made by the target generates 1 SB (Hearts) as the effort to maintain falsehoods under the Witness's gaze creates internal discord. \\
\textbf{Push It:} You may designate one specific, complex question about the target's motivations, fears, or hidden loyalties. If you successfully use Sway or Insight against them this scene, you automatically learn the answer to that question. The intense scrutiny marks 1 SB (Spades) for you as the Witness's attention lingers. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 6 segments. \\
\emph{Note:} The Witness strips away the veils of the heart. % Revised/replaced Burden of Memory

\paragraph{Rite of the Final Reckoning [OMEN] (High, 13 XP)} \emph{Scene; Zone; No.}
\textbf{Materials:} A formally called gathering (court, council, family meeting) within the consecrated zone. \\
\textbf{Effect:} All present must speak their greatest debt, wrongdoing, or hidden truth related to the gathering's purpose. Those who lie or withhold suffer Harm 2 (Stress/Reputation). Truth-tellers gain +2 dice to social actions for the remainder of the scene within the zone. \\
\textbf{Push It:} The truth becomes inescapable - even indirect lies or evasions related to the core topic suffer the Harm 2 penalty. The absolute nature of the revelation creates 2 SB (Diamonds) as the disruption to fates and secrets resonates. \\
\emph{Requires: Familiar + Codex + Tier III \ (\textit{Invoke:} \textbf{2 Boons}).} \\
\emph{Obligation:} 7 segments. \\
\emph{Note:} The Witness calls the ultimate accounting. % New High Rite to replace the weak Burden of Memory
