% =========================================================
% END OF APPENDICES (CHAPTER 12)
% =========================================================

\clearpage
\chapter{Using Expansions: A Guide for the Overwhelmed GM}
\label{ch:using-expansions}
\index{expansions}

\begin{quote}
\small
\emph{``I bought every expansion. Then I sat at my table with thirty-seven PDFs and a blank look. The players were waiting. The beer was warm. I learned that night: you don't use expansions. You borrow from them.''}
--- The Gray Wanderer, from a letter found under a stack of printouts
\end{quote}

\section{The First Rule: You Own the Expansions, They Don't Own You}

Every expansion in Fate's Edge is a toolbox, not a syllabus. You are never required to use a single rule, timer, or cultural detail from any expansion. The core game runs perfectly well on its own. Expansions exist to \emph{solve specific problems} or \emph{add specific flavors} when you need them.

\subsection{The Three Questions to Ask Before Opening Any Expansion}

Before you reach for a PDF, ask yourself:

\begin{enumerate}
\item \textbf{What problem am I solving?} (e.g., ``combat feels same-y'', ``my players want more political intrigue'', ``we're going to sea'')
\item \textbf{What is the smallest piece of this expansion that solves it?} (e.g., ``just the Seven Bell Court judging table'', ``just the travel timers'', ``just the Lethai-ar witness knots'')
\item \textbf{Can I explain that piece to my players in one minute?} If not, pick a smaller piece.
\end{enumerate}

\subsection{The 10\% Rule}

A typical Fate's Edge expansion is 30-100 pages. In a single session, you will use \textbf{at most 10\% of any expansion}. The other 90\% is background, lore, alternative systems, and optional depth. You are not expected to master all of it.

\section{A Taxonomy of Expansions (What They're Actually For)}

Use this table to quickly identify which expansion to reach for. Expansions are grouped by \textbf{purpose}, not by title.

\begin{longtable}{p{3.5cm}|p{3.5cm}|p{6cm}}
\hline
\textbf{If you want...} & \textbf{Use expansions in this family} & \textbf{Example of a single mechanic to steal} \\
\hline
Honorable duels, tournaments, social combat & Seven Bell Court, Vhasia: The Knave's Regret & The Form/Spirit/Intent judging table (3 rows, no dice) \\
\hline
New martial options (no magic) & The Way of the Warrior, Ykrul: Iron \& Blood & A single talent chain (e.g., Sentinel, Bladestorm, Pathfinder) \\
\hline
Cultural depth for a specific people & Eastern Realms, Peoples and Cultures, The Frost Ledger & One etiquette hook and one String (e.g., ``Aelinnel never ring the ninth bell'') \\
\hline
Investigation & The Salt Ledger, The Wandering Ledger, Saikou Ira's Compendium & One clue type (e.g., ``a silver coin in the mouth = ferryman's toll'') \\
\hline
Sea travel, naval combat, port intrigue & The Amaranthine Sea & Travel timers [4-8] and the Inspection procedure \\
\hline
Folk horror, psychological tension & Horror Campaigns, The Frost Ledger, Mistslands expansion & The Dread Timer (4-8 segments) \\
\hline
Witchcraft, curses, bargains & Witches of Fate's Edge, Daughters of the Unbroken Cord & One working type (e.g., Chord Working, Promise Timer , Cord of Flesh) \\
\hline
Patron-heavy campaigns & The Gray Wanderer's Grimoire, Malachai, The Chained Angel & One new Patron with its Rites and Obligation flavor \\
\hline
Dungeon crawling, ancient lairs & Dragon's Lair, The Book of Air, Mist, Alder, Thorn, and Mirror & Lair moves (3-4 per scene) \\
\hline
Political intrigue & Political Intrigue framework, Vhasia, Viterra expansions & Faction timers (Influence/Stability/Exposure) \\
\hline
Caravans, travel, trade & Caravans: The Way of Silk, The Wandering Ledger & Travel intent dial (Speed/Stealth/Show/Tribute) \\
\hline
\end{longtable}

\section{The Hybrid Adventure: How to Cook with Leftovers}

The most successful way to use expansions is to build a \textbf{hybrid adventure}: one core story (often from a free adventure or your own imagination) with mechanics borrowed from 2-4 expansions.

\subsection{Example: The Tournament of the Broken Crown}

The GM wanted:
\begin{itemize}
\item A tournament (so players could fight without dying).
\item A murder mystery (so non-combat characters had something to do).
\item Cultural flavor (so the world felt lived-in).
\end{itemize}

They borrowed:
\begin{itemize}
\item \textbf{Judging table} from \textit{Seven Bell Court} (3 rows: Form, Spirit, Intent).
\item \textbf{Unbuckled gorget clue} from \textit{Vhasia: The Knave's Regret} (mercy as a theme).
\item \textbf{Bell-tending} and \textbf{witness knots} from \textit{Eastern Realms} (investigation hooks).
\item \textbf{Martial talents} from \textit{The Way of the Warrior} (Disciplined Body, Shield Bearer, Flurry).
\end{itemize}

\textbf{Prep time:} 90 minutes. \textbf{Pages read:} \(\sim 15\) total across four expansions. \textbf{Result:} A 3-hour session that felt fresh, deep, and manageable.

\subsection{The Hybrid Recipe}

\begin{enumerate}
\item Pick a \textbf{core activity} (tournament, heist, travel, investigation).
\item Identify \textbf{one gap} in the core rules that an expansion fills (e.g., ``combat lacks stakes beyond HP'' \(\rightarrow\) Seven Bell Court).
\item Read \textbf{only the first 5-10 pages} of that expansion until you find a mechanic you understand.
\item Add \textbf{one cultural hook} from a lore expansion (e.g., ``in this city, knights unbuckle their gorgets to show mercy'').
\item Add \textbf{one talent} from a player-options expansion (e.g., give an NPC a talent from \textit{The Way of the Warrior}).
\item Run the session. \textbf{Ignore every other rule} from the expansions.
\end{enumerate}

\section{Managing Player Expectations}

When players know you own expansions, they will ask to use things. This is a gift, not a burden.

\subsection{The ``Yes, But Later'' Rule}

When a player asks ``Can I use the Shadow Step talent from \textit{Witches of Fate's Edge}?''

\begin{itemize}
\item \textbf{Yes:} ``Yes, but you'll need to earn it through roleplay. Find a hag or a Lethai-ar teacher in the next town.''
\item \textbf{No, but:} ``No, but you can take a different talent from the core book that works similarly, and we'll reskin it.''
\item \textbf{Later:} ``Not this session, but tell me about it during downtime and we'll build it into your character's arc.''
\end{itemize}

\subsection{The Expansion Session Zero}

At the start of a new campaign or arc, ask your players:

\begin{enumerate}
\item ``Which expansions have you read?'' (Be honest; they may have read none.)
\item ``What's one thing from an expansion you'd love to see in play?'' (Write it down.)
\item ``What's one thing you absolutely don't want?'' (Lines and veils for expansion content.)
\end{enumerate}

Then pick \textbf{one expansion} to feature prominently in the next 2-3 sessions. Rotate.

\section{Cross-Expansion Synergy (When Two Books Work Better Than One)}

Some expansions are designed to work together. Others accidentally complement each other. Here are known synergies:

\begin{itemize}
\item \textbf{Seven Bell Court + Vhasia: The Knave's Regret} = duels that are judged on honor, not just damage. The ``unbuckled gorget'' becomes a mechanical choice (lower Defense, higher Spirit points).
\item \textbf{The Amaranthine Sea + The Salt Ledger} = sea travel with port mysteries. Use the travel timers from \textit{Sea} and the lighthouse/spirit stories from \textit{Salt} as random encounters.
\item \textbf{Witches of Fate's Edge + Horror Campaigns} = curse-based horror. Use the Dread Timer and the Cord mechanics together; each curse failure ticks the Dread Timer .
\item \textbf{Eastern Realms + The Way of the Warrior} = culturally-infused martial characters. Let a Lethai-al samurai take Bushido-style talents from \textit{Warrior} and reflavor them as ``the Way of the Leaf.''
\end{itemize}

\subsection{Antisynergy: When Not to Mix}

Some expansions actively fight each other. Avoid:

\begin{itemize}
\item \textbf{Too many timers.} \textit{Horror Campaigns} (Dread Timer , Reality Fracture) + \textit{Political Intrigue} (Mandate, Crisis) + \textit{The Amaranthine Sea} (Weather Window, Patrol Net) = 6+ timers. Players cannot track that many. Pick 2-3 maximum per session.
\item \textbf{Contradictory cultural lore.} The \textit{Eastern Realms} and \textit{The Salt Ledger} have different versions of Aelinnel bell-customs. Choose one interpretation for your campaign and stick to it. Tell players ``we're using the \textit{Eastern Realms} version.''
\end{itemize}

\section{Practical Tools: The Expansion Integration Checklist}

Before a session where you plan to use expansion content, run through this checklist.

\subsection{Pre-Session (30-60 minutes)}

\begin{enumerate}
\item \textbf{Select 1-3 expansions} to draw from. Write their names on an index card.
\item \textbf{Extract 1-3 mechanics} from each expansion. Write each mechanic as a single sentence on the card.
\item \textbf{Prepare one page of reference.} Print the relevant table (e.g., Seven Bell Court judging, travel timer sizes, talent list). Do not print the whole expansion.
\item \textbf{Note page numbers.} If you must look something up, know where it is.
\item \textbf{Check for conflicts.} Do any of your chosen mechanics use the same resource (e.g., both use Boons in contradictory ways)? If yes, pick one.
\end{enumerate}

\subsection{During Session (At the Table)}

\begin{enumerate}
\item \textbf{Announce the new mechanic.} ``Tonight we're using the Seven Bell Court. Your duels will be judged on Form, Spirit, and Intent.''
\item \textbf{Demonstrate once.} Run a quick mock duel or skill challenge to show how it works.
\item \textbf{Let players opt out.} If someone doesn't want to use a new subsystem, let them use the core rules instead.
\item \textbf{Keep a timer.} If a new mechanic slows play for more than 2 minutes, simplify it or drop it for the scene.
\end{enumerate}

\subsection{Post-Session (5 minutes)}

\begin{enumerate}
\item \textbf{Ask:} ``What expansion mechanic worked? What didn't?''
\item \textbf{Note:} Write down which pages you actually used. Toss the rest.
\item \textbf{Decide:} Will you use this expansion again? If yes, put the relevant page in your GM binder. If no, archive it.
\end{enumerate}

\section{DM's Cheat Sheet: Quick Reference by Page Count}

If you have limited time before a session, use this triage guide.

\begin{longtable}{p{4cm}|p{4cm}|p{4cm}}
\hline
\textbf{Expansion} & \textbf{Read these pages} & \textbf{Skip these pages} \\
\hline
Seven Bell Court & 5-7 (judging table, Honor/Pragmatism) & Tournament logistics, NPC champion statblocks \\
\hline
Vhasia: The Knave's Regret & 1-3 (the opening fiction), 25-27 (talents) & Full canton descriptions, duel history \\
\hline
Way of the Warrior & 31 (talent quick reference) & Cultural warrior variants (unless a player is one) \\
\hline
Eastern Realms & 13-14 (bell-tending, witness knots) & Travel decks (use only when traveling) \\
\hline
Witches of Fate's Edge & i-iv (first principles), 89-91 (Threshold working) & Deep hag lore, full talent trees \\
\hline
The Amaranthine Sea & 9-11 (travel timers, inspection) & Fleet combat rules, ship statblocks \\
\hline
Horror Campaigns & 5-6 (Dread Timer ), 8-9 (breaking points) & Extended campaign timers (use only for long arcs) \\
\hline
Dragon's Lair & 3-5 (lair generator), 21 (dragon moves) & Hoard generators, dragon stats (improvise Tier) \\
\hline
The Salt Ledger & Any single story (use as adventure seed) & The rest of the stories (save for later) \\
\hline
\end{longtable}

\section*{Magic Adjudication Quick Reference (GM)}
\label{sec:gm-magic-cheat}

\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=Default Magic DV,breakable]
\centering
\begin{tabular}{ll}
\textbf{Effect scope} & \textbf{DV} \\
Cantrip (1 tag) & 2 \\
Combat spell (2 tags) & 3 \\
Ritual (small, 3 tags) & 4 \\
Ritual (large, dangerous tags) & 5+ \\
\end{tabular}
\end{tcolorbox}

\subsection*{Free Casting (TAGS)}
\[
DV = 1 + \text{\#tags} \quad(\text{dangerous tags: } +2\text{ instead of }+1)
\]
\emph{Dangerous tags:} Leap, Fold, Gate, Gravity, Create, Summon, Transmute, Animate, Transform, Dominate, Rewind, Resurrect.

\paragraph{Backlash:} Spend SB from 1s. Severity by Position:
\begin{itemize}
  \item Dominant: minor (noise, Fatigue 1)
  \item Controlled: moderate (Harm 1, Condition)
  \item Desperate: severe (Harm 2, scene shift)
\end{itemize}

\subsection*{Rites (Runekeepers \& Invokers)}
\[
DV = \max(\text{ObligationCost} - \text{Spirit},\; \text{Tier: Low=1, Standard=2, High=3})
\]
\paragraph{Obligation overflow:} Exceed Capacity (Spirit+Presence) \(\rightarrow\) mark 1 Fatigue per segment. Exceed 2\(\times\) Capacity \(\rightarrow\) +1 Harm, Patron intrusion.

\subsection*{Healing Magic}
\begin{itemize}
  \item \textbf{[HEAL] only:} Restores 1 Fatigue. DV 2. 1 action.
  \item \textbf{[HEAL] + another tag (Purify, Strengthen, Cleanse):} Restores 1 Harm. DV = 2 + number of tags, plus \textbf{+2 surcharge}. Requires two actions unless Push.
  \item \textbf{Ritual healing (Harm 2+):} Requires Extended ritual (timer), rare components.
\end{itemize}

\subsection*{Summoning}
\begin{itemize}
  \item \textbf{Call:} 1 action, spirit manifests.
  \item \textbf{Bind:} 1 Boon or 1 Fatigue.
  \item \textbf{Leash capacity:} Cap + Spirit segments. Tick when: spirit takes Harm, acts against nature, crosses a [WARD], summoner splits focus, rival contests control.
  \item \textbf{Boon Finesse:} Once/round, spend 1 Boon to clear 1 Leash tick.
  \item \textbf{Leash full:} Spirit acts once to its nature, then departs (or turns hostile).
\end{itemize}

\subsection*{Cantors (Songs)}
\begin{itemize}
  \item \textbf{Sing:} 1 action, resolves next turn. Roll Lore + Performance (DV 2--3).
  \item \textbf{Push:} Resolve immediately, mark 1 Fatigue, advance Corruption Timer by 1.
  \item \textbf{Corruption Timer :} Size = Spirit. When full \(\rightarrow\) bloom (gain benefit \& burden from last patron), reset to Tier (min 1).
  \item \textbf{Resonant Rite:} Performance with 10+ witnesses \(\rightarrow\) may found Chorus (minor follower benefit, Obligation to patron each session).
\end{itemize}

\subsection*{Patron Intrusions (when Obligation overflows or GM spends 2+ SB)}
\begin{itemize}
  \item Demand a quest (``Find my stolen symbol.'')
  \item Impose a cost (``Your next Rite costs double.'')
  \item Send an omen (``Candles burn blue; you feel watched.'')
  \item Summon a rival servant (``Another follower of Malachai is hunting you.'')
\end{itemize}

\subsection*{Common Magic Rulings}
\begin{itemize}
  \item \textbf{Countermagic:} [COUNTER] tag vs. DV = spell's DV. Success cancels.
  \item \textbf{Dispel:} [DISPEL] vs. DV = effect's Tier (1--5). On hit, effect ends.
  \item \textbf{Wards:} [WARD] vs. Outsider. DV = target's Cap. Crosser adds DV to Leash.
  \item \textbf{Line of sight:} Required for most spells; without it +1 DV.
\end{itemize}

\section{Adversary Templates}

\subsection{Generic Adversaries}
\label{subsec:generic-adversaries}
Use these for filler encounters. No special traits---just core stats. Scale by adjusting Threat Level (see Table~\ref{tab:threat-scaling}). All use standard attack/defense as per Chapter~\ref{ch:gm-core}.

\begin{itemize}
    \item \textbf{Bandit Rabble} (Threat Level 1) \\
    \textit{Stats:} Body 2, Wits 1, Spirit 1, Presence 1 \quad | \quad \textit{Skills:} Melee 1, Athletics 1 \\
    \textit{Flavor:} ``Desperate locals with rusty blades. They flee if half their number falls.'' \\
    \textit{Use:} Random encounters, tavern brawls, distractions.

    \item \textbf{City Watch Recruit} (Threat Level 2) \\
    \textit{Stats:} Body 2, Wits 1, Spirit 1, Presence 2 \quad | \quad \textit{Skills:} Melee 1, Command 1 \\
    \textit{Flavor:} ``Poorly trained but eager to prove themselves. Will call for backup if overwhelmed (alarm raises a 4-segment timer).'' \\
    \textit{Use:} Urban patrols, gate guards, low-level enforcement.

    \item \textbf{Cult Novice} (Threat Level 1) \\
    \textit{Stats:} Body 1, Wits 1, Spirit 2, Presence 1 \quad | \quad \textit{Skills:} Lore 1 (cult-specific), Sway 1 \\
    \textit{Flavor:} ``True believer, poorly trained. Will gladly die for the cause but breaks if shown proof their faith is a lie.'' \\
    \textit{Use:} Guarding rituals, spreading propaganda, cannon fodder. On a Miss when trying to recruit or intimidate, they may flee or surrender (GM discretion).

    \item \textbf{Street Thug} (Threat Level 2) \\
    \textit{Stats:} Body 3, Wits 1, Spirit 1, Presence 1 \quad | \quad \textit{Skills:} Brawl 2, Intimidation (Presence+Brawl) \\
    \textit{Flavor:} ``Enforcer for a local gang. Favors cheap shots and overwhelming numbers.'' \\
    \textit{Use:} Alley ambushes, protection rackets, gang warfare.

    \item \textbf{Guard Sergeant} (Threat Level 3) \\
    \textit{Stats:} Body 3, Wits 2, Spirit 2, Presence 2 \quad | \quad \textit{Skills:} Melee 2, Command 1, Athletics 1 \\
    \textit{Flavor:} ``Veteran who has seen it all. Leads from the front and shouts orders.'' \\
    \textit{Use:} Leader of a watch patrol, military outpost, or noble's bodyguard.
\end{itemize}

\begin{table}[htbp]
\centering
\caption{Threat Level Scaling}
\label{tab:threat-scaling}
\begin{tabular}{lcll}
\toprule
\textbf{Threat Level} & \textbf{Stat Adjustment} & \textbf{Typical Use} & \textbf{PC Tier Match} \\
\midrule
0 & --1 to all stats & Non-threatening flavor & Below PC tier \\
1 & Base stats (as written) & Standard minions & PC tier --1 \\
2 & +1 to all stats & Tougher minions / elite & PC tier \\
3 & +2 to all stats & Mini-boss equivalent & PC tier +1 \\
4+ & +3 to all stats \& add one archetype trait & Major boss & PC tier +2 \\
\bottomrule
\end{tabular}
\end{table}
\par\smallskip\textit{Adjust all Attributes and Skills uniformly. For Threat Level 0, treat as flavor only (no combat threat). For Threat Level 3+, consider giving one simple trait from the archetypes below.}

\subsection{Archetypal Adversaries}
\label{subsec:archetypal-adversaries}
Each has one thematic mechanic tied to existing rules. Use as recurring threats, scene-setters, or mini-bosses. Scale via Threat Level table.

\subsubsection{The Iron Patriarch / Matriarch}
\textit{Archetype:} Ruthless leader of a family, guild, or corporation. Values legacy and control above all.\\
\textit{Stats (Threat Level 2):} Body 2, Wits 3, Spirit 2, Presence 3 \quad | \quad \textit{Skills:} Sway 2, Command 1, Lore 1 (family/guild)\\
\textbf{Thematic Mechanic:} \textit{Legacy's Weight} \\
When the Patriarch/Matriarch would take Harm, they may immediately order a loyal subordinate (if present) to take the blow. Roll Wits + Command \textit{vs. DV 2}. On success, the subordinate takes 1 Harm instead (armor applies normally). On failure, the Patriarch takes the Harm but gains +1 die on their next action (resolve fueled by desperation).\\
\textit{Flavor Hook:} ``They don't fear death---they fear their life's work dying with them. Every blow you strike feels like it's aimed not at them, but at the empire they bled to build.''\\
\textit{Rules Tie-In:} Uses opposed rolls (see Chapter~\ref{ch:gm-core}) and Harm (see Chapter~\ref{ch:gm-core}).

\subsubsection{The Hollow Prophet}
\textit{Archetype:} Charismatic leader of a movement built on a lie (or partial truth). Believes their own deception.\\
\textit{Stats (Threat Level 2):} Body 1, Wits 2, Spirit 3, Presence 3 \quad | \quad \textit{Skills:} Sway 2, Lore 1 (doctrine), Deception 1\\
\textbf{Thematic Mechanic:} \textit{Faith's Fragility} \\
When the Hollow Prophet successfully uses Sway to convert a bystander or reinforce a follower's faith, they generate \textbf{1 extra Story Beat} (beyond any generated by 1s). However, if the Sway attempt \textbf{fails} (Miss), they gain a \textit{Despair} Condition: --1 die to all Spirit and Wits rolls until they successfully persuade someone or gain 1 Boon.\\
\textit{Flavor Hook:} ``Their eyes glow with conviction as they speak---but when you catch their reflection in a dark window, you see only exhaustion and doubt. The lie sustains them... but how long until it cracks?''\\
\textit{Rules Tie-In:} Uses Story Beats (\autoref{sec:story-beats-backlash}) and Conditions (see Chapter~\ref{ch:gm-core}).

\subsubsection{The Gearsmith}
\textit{Archetype:} Obsessed tinkerer, weaponsmith, or arcane engineer. Sees people as problems to solve.\\
\textit{Stats (Threat Level 2):} Body 2, Wits 3, Spirit 1, Presence 2 \quad | \quad \textit{Skills:} Craft 2, Mechanics 1, Brawl 1\\
\textbf{Thematic Mechanic:} \textit{Improvised Countermove (Cost: 1 Boon)} \\
When the Gearsmith is targeted by an action (attack, skill use, etc.), the GM may spend 1 Story Beat (or 1 Boon from the GM's pool -- see \autoref{sec:boons-gm}) to invent a reactive device on the spot. The device negates the incoming action and imposes one relevant Condition (e.g., \textit{Entangled}, \textit{Blinded}, \textit{Shaken}) on the attacker. The device is single-use and narratively described (GM improvises based on the scene). This ability can be used at most once per round.\\
\textit{Flavor Hook:} ``They don't dodge your sword---they're already calculating how to turn its momentum against you. That clang wasn't a block; it was the sound of a spring-loading.''\\
\textit{Rules Tie-In:} Uses Story Beats (\autoref{sec:story-beats-backlash}) and Conditions (see Chapter~\ref{ch:gm-core}).

\subsubsection{The Iron-Bound Rival (PC Mirror)}
\textit{Archetype:} Former ally, rival adventurer, or ideological opponent who mirrors the PCs' path.\\
\textit{Stats:} Match the PC's highest two Attributes at their current rating. Match the PC's two highest Skills at their current rating. Other stats default to 1 or 0 as appropriate.\\
\textbf{Thematic Mechanic:} \textit{Anticipate Move} \\
When a PC declares an action targeting the Rival, the Rival gains \textbf{+1 die} to any defensive or opposing roll (e.g., dodging an attack, resisting Sway, countering a skill use). This bonus does not stack if multiple PCs act against the Rival in the same exchange.\\
\textit{Flavor Hook:} ``They move not like a fighter, but like a shadow of your own choices---always one step ahead of where you \textit{think} you're going, but never where you \textit{are}.''\\
\textit{Rules Tie-In:} Uses opposed rolls (see Chapter~\ref{ch:gm-core}) and mirrors PC progression.

\subsubsection{The Weeping Martyr}
\textit{Archetype:} Someone who has lost everything and fights with suicidal abandon.\\
\textit{Stats (Threat Level 2):} Body 3, Wits 1, Spirit 3, Presence 1 \quad | \quad \textit{Skills:} Melee 2, Endurance 2, Intimidation (Presence+Brawl)\\
\textbf{Thematic Mechanic:} \textit{Blood Price} \\
When the Martyr takes Harm, they immediately gain +1 die on their next attack or physical action (pain fuels fury). Additionally, if they are reduced to Harm 3 (incapacitated), they may make one final attack as a free action before falling. This attack ignores all Position penalties and deals +1 Harm.\\
\textit{Flavor Hook:} ``They don't even try to block your strike---they lean into it, grinning through bloody teeth. Every wound is another coin in a purse they'll never spend.''\\
\textit{Rules Tie-In:} Uses Harm levels (see Chapter~\ref{ch:gm-core}) and the ``roller-coaster'' effect (see \texttt{[MISSING REF: sec:roller-coaster]}).

\subsubsection{The Shadow Broker}
\textit{Archetype:} Information dealer, blackmailer, or spymaster who never fights fair.\\
\textit{Stats (Threat Level 2):} Body 1, Wits 4, Spirit 2, Presence 2 \quad | \quad \textit{Skills:} Subterfuge 3, Insight 2, Streetwise 2\\
\textbf{Thematic Mechanic:} \textit{I Know a Guy (Cost: 1 Story Beat)} \\
Once per scene, the Shadow Broker may spend 1 Story Beat to declare that they have prepared for the current situation. They produce a minor but useful asset (e.g., a hidden exit, a bribed guard, a forged document) that grants the party (or their own allies) +1 die or improves Position by one step on a single roll. The GM narrates the asset and its limitations.\\
\textit{Flavor Hook:} ``You never see them draw a weapon. You only see the trap they set three scenes ago, the witness they paid off, the escape route you didn't know existed.''\\
\textit{Rules Tie-In:} Uses Story Beats (\autoref{sec:story-beats-backlash}) and Assets (\autoref{sec:followers-assets-gm}). The Broker rarely engages in direct combat; instead, they create advantages.

\subsection{Using These Templates in Play}
\begin{itemize}
    \item \textbf{Generic foes}: Drop in instantly. Adjust Threat Level via Table~\ref{tab:threat-scaling}. No prep beyond naming them.
    \item \textbf{Archetypal foes}: Pick one matching your scene's theme. Note their one mechanic. Use the \textit{Flavor Hook} to guide narration.
    \item \textbf{Scaling}: If PCs are Tier III, treat Threat Level 2 as baseline. For major encounters, use Threat Level 3 or 4 archetypes.
    \item \textbf{No extra bookkeeping}: Mechanics reference existing rules---just apply them as written. The mechanic is the \textit{only} thing to remember beyond core stats.
\end{itemize}

\noindent
\rule{\linewidth}{0.5pt}
\small
\emph{Design Philosophy: Generic foes exist to threaten through numbers and positioning (using core rules only). Archetypal foes add narrative depth via one self-contained mechanic that reflects their core theme---never requiring tracking beyond the encounter. All mechanics are verifiable in the core rules (see citations in the text).}
