\clearpage
\section*{Appendix C: Deck and Magic Reference}
\addcontentsline{toc}{section}{Appendix C: Deck and Magic Reference}
\label{app:deck-magic}

\begin{tcolorbox}[title=Deck Types and Meanings, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{decks}
\begin{description}
  \item[Travel Decks (regional, 52-card)] Spade = Place, Heart = Actor, Club = Pressure, Diamond = Leverage.
  \item[Deck of Consequences (scene drama)] Hearts = Social/Emotional, Spades = Harm/Escalation, Clubs = Material Cost, Diamonds = Magical/Spiritual.
\end{description}
\textbf{Important:} Never mix suit meanings across different deck types.
\end{tcolorbox}

\begin{tcolorbox}[title=Deck Usage Procedure, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{deck of consequences!procedure}
After a roll that generates Story Beats:
\begin{enumerate}
  \item \textbf{Direct Spend:} Translate SB into immediate consequences or timer ticks (see Appendix A).
  \item \textbf{Optional Deck Draw:} Draw up to \(\min(\text{SB},3)\) cards and synthesize a single narrative twist guided by the suits and the highest rank.
\end{enumerate}
\end{tcolorbox}

\begin{tcolorbox}[title=Rank Severity Guide, colback=white!97!gray, colframe=black!80!gray,breakable]
\begin{description}
  \item[Ace--3] Minor inconvenience or flavor.
  \item[4--6] Moderate setback.
  \item[7--9] Significant consequence.
  \item[10--King] Major fallout with lasting effects.
\end{description}
\end{tcolorbox}

\begin{longtable}{lp{10cm}}
\caption{Deck of Consequences Interpretation} \label{tab:consequences-interp} \\
\toprule
\textbf{Suit} & \textbf{Interpretation} \\
\midrule
\endfirsthead
\multicolumn{2}{c}%
{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Suit} & \textbf{Interpretation} \\
\midrule
\endhead
\bottomrule
\endfoot
Hearts (Emotional/Social) & Ace--3: awkward social moment; 4--6: relationship strain; 7--9: betrayal; 10--King: exile or shattered alliance. \\
Spades (Harm/Escalation) & Ace--3: bruise or Fatigue; 4--6: wound or damaged gear; 7--9: severe injury; 10--King: death or permanent loss. \\
Clubs (Material/Cost) & Ace--3: minor resource loss; 4--6: gear failure or debt; 7--9: major asset lost; 10--King: total ruin. \\
Diamonds (Magical/Spiritual) & Ace--3: omen or whisper; 4--6: curse or spirit appears; 7--9: arcane backlash; 10--King: reality bends temporarily. \\
\end{longtable}
\index{deck of consequences!interpretation}

\begin{longtable}{llll}
\caption{Magic Paths Comparison} \label{tab:magic-paths} \\
\toprule
\textbf{Path} & \textbf{Requirements} & \textbf{Key Feature} & \textbf{Risk Type} \\
\midrule
\endfirsthead
\multicolumn{4}{c}%
{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Path} & \textbf{Requirements} & \textbf{Key Feature} & \textbf{Risk Type} \\
\midrule
\endhead
\bottomrule
\endfoot
Caster (Freeform) & Spellcraft (6 XP) & Improvisational spellcasting & Backlash (\ref{subsec:elemental-backlash}) \\
Runekeeper (Rites) & Thiasos (2 XP) + Codex (4 XP) & Structured Patron Rites & Obligation (\ref{sec:obligation-corruption-gm}) \\
Invoker (Symbols) & Patron's Symbol (4 XP each) & Ritual precision, emergency casting & Symbol compromise (\ref{subsec:corruption-blooms}) \\
\end{longtable}
\index{magic!paths comparison}

\begin{tcolorbox}[title=DV Reference for Rites, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{rites!difficulty value}
\[
\mathrm{DV} = \max(\text{ObligationCost} - \text{Spirit},\; \text{Tier})
\]

\begin{tabular}{c|cccc|c|c|c}
Obligation & Spirit 0 & Spirit 1 & Spirit 2 & Spirit 3--4 & Tier 1 & Tier 2 & Tier 3 \\
\hline
1 & 1 & 1 & 1 & 1 & 1 & 2 & 3 \\
2 & 2 & 1 & 1 & 1 & 1 & 2 & 3 \\
3 & 3 & 2 & 1 & 1 & 1 & 2 & 3 \\
4 & 4 & 3 & 2 & 1 & 1 & 2 & 3 \\
5 & 5 & 4 & 3 & 2 & 1 & 2 & 3 \\
6 & 6 & 5 & 4 & 3 & 1 & 2 & 3 \\
7 & 7 & 6 & 5 & 4 & 1 & 2 & 3 \\
\end{tabular}
\end{tcolorbox}

\begin{tcolorbox}[title=Casting Loop Summary, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{free casting!procedure}
\begin{enumerate}
  \item \textbf{Channel:} Wits + Arcana to gather magical potential.
  \item \textbf{Weave:} Wits + Art (or appropriate magical skill) to shape the effect.
  \item \textbf{Backlash:} The GM spends SB generated during the roll to impose thematic consequences (see \ref{subsec:elemental-backlash}).
\end{enumerate}
\end{tcolorbox}

\begin{tcolorbox}[title=Eight Elements of Magic, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{elements of magic}
\begin{description}
  \item[Earth] Solidity, stability, foundation.
  \item[Fire] Energy, transformation, destruction.
  \item[Air] Movement, speed, freedom.
  \item[Water] Fluidity, healing, adaptability.
  \item[Fate] Destiny, inevitability, causality.
  \item[Life] Vitality, creation, growth.
  \item[Luck] Chance, unpredictability.
  \item[Death/Dreams] Endings, thresholds, the subconscious.
\end{description}
\end{tcolorbox}
