\clearpage
\section*{Appendix B: Optional Rules and Tactical Tools}
\addcontentsline{toc}{section}{Appendix B: Optional Rules and Tactical Tools}
\label{app:optional-rules}

\begin{tcolorbox}[title=Grid-Based Combat (Optional), colback=white!97!gray, colframe=black!80!gray,breakable]
\index{combat!grid}
Use a square or hex grid with Zones of Control (ZoC).

\textbf{ZoC rules:}
\begin{itemize}
  \item Entering an enemy's Zone of Control ends your movement -- you are engaged.
  \item You cannot move through an enemy's ZoC.
  \item Leaving a ZoC requires the Disengage action (DV 4--6) or spending 1 Boon.
  \item Flank: \(+1\) die to attack. Rear: \(+1\) die and \(+1\) Effect.
\end{itemize}
\textbf{Push-Through:} Roll \(\text{Body}+\text{Athletics}\) vs. DV \(= 2 + \text{enemy Tier}\). Choose a cost on success: Position worsens by one step, --1 die on the action, or the enemy gets a free opportunity attack.
\end{tcolorbox}

\begin{tcolorbox}[title=Detailed Warfare (Mass Combat), colback=white!97!gray, colframe=black!80!gray,breakable]
\index{combat!mass}
Treat an army as a high-Cap Follower (see \ref{sec:followers-assets-gm}) with its own timers:
\begin{itemize}
  \item \textbf{Supply Timer } (6--8 segments): When empty, the army suffers severe penalties.
  \item \textbf{Morale Timer } (6--8 segments): When full, the army routs.
  \item \textbf{Command rolls} (Presence + Command or Wits + Tactics) advance objective timers (e.g., \emph{Break Their Line} [6]).
  \item \textbf{Battle's Edge:} Story Beats generated by command rolls can be spent on fog-of-war complications -- false orders, broken supply wagons, or a messenger's death.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Hex-Based Exploration, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{exploration!hex}
Map a region in 6-mile hexes. Entering a hex requires a \(\text{Wits}+\text{Survival}\) roll (DV set by terrain). Outcomes determine discovery (find a point of interest), danger (ambush, hazard), or becoming lost (add a timer segment to find the way out).
\end{tcolorbox}

\begin{tcolorbox}[title=Miniatures Mode --- Speed Defaults, colback=white!98!gray, colframe=black!80!gray,breakable]
\index{miniatures}
\textbf{Base DV:} \(\mathrm{DV}=\mathrm{Tier}+2+\text{Keywords}\) (elevation \(+1\), \texttt{[WARD]} \(+1\), Disengage base DV 4).\\
\textbf{Critical Hit (10):} Bump Position one step; if already Dominant, Push or Pull the target 1 hex \emph{or} gain \(+1\) automatic success.\\
\textbf{Zone of Control (ZoC):} Entering or leaving a hex adjacent to an enemy provokes one Reaction (Free Strike or Shove 1 hex).\\
\textbf{Heat:} On a critical hit, the GM may spend 1 Heat (a resource tracked per scene) to degrade Position or trigger a terrain hazard.
\end{tcolorbox}

\begin{tcolorbox}[title=Hex Keywords, colback=white!98!gray, colframe=black!80!gray,breakable]
\index{hex keywords}
\textbf{Difficult:} Costs 2 movement points per hex. 
\textbf{Elevation:} Attacks from lower ground suffer \(+1\) DV. 
\textbf{ZoC:} Provokes a Reaction on crossing.\\
\textbf{Altar \texttt{[WARD]}:} --1 die to attacks (or attacker may spend 2 Fatigue to ignore).\\
\textbf{Incorporeal:} Ignores Difficult terrain; can pass through occupied hexes but cannot end movement in one.
\end{tcolorbox}

\subsection*{Sample NPCs}
\label{sec:sample-npcs}

\begin{tcolorbox}[title=Encounters (Tier I--II), colback=white!97!gray,breakable]
\begin{itemize}
  \item \textbf{Bandit Skirmisher:} Body 2, Wits 2. Melee 2, Stealth 1.
  \item \textbf{Ykrul Rider:} Body 4, Wits 3. Riding 3, Melee 3.
  \item \textbf{Street Bravo:} Presence 3, Body 2. Dueling 3.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Foils \& Rivals (Tier II--III), colback=white!97!gray,breakable]
\begin{itemize}
  \item \textbf{Ambitious Scribe:} Wits 3, Presence 3. Intrigue 3, Lore 2.
  \item \textbf{Mercenary Captain:} Body 4, Spirit 3. Command 3, Melee 4.
  \item \textbf{Flame Preacher:} Presence 4, Spirit 3. Oratory 4, Faith 3.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Prestige NPCs (Tier IV--V), colback=white!97!gray,breakable]
\begin{itemize}
  \item \textbf{High Elf Loremaster:} Wits 5, Spirit 4. Lore 5, Arcana 4.
  \item \textbf{Dwarven Forge-Patriarch:} Body 5, Spirit 4. Craft 5, Command 4.
  \item \textbf{Ykrul Warglord:} Body 5, Presence 4. Command 4, Melee 5.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Common Rolls (GM Screen Reference), colback=white!97!gray, colframe=black!80!gray,breakable]
\index{skill rolls!common examples}
\begin{multicols}{2}
\textbf{Athletics:} climb (\emph{Body}+Athletics), time a leap (\emph{Wits}+Athletics)\\
\textbf{Stealth:} shadow patrol (\emph{Wits}+Stealth), hold still (\emph{Spirit}+Stealth)\\
\textbf{Endurance:} resist cold (\emph{Spirit}+Endurance), push through pain (\emph{Spirit}+Endurance)\\
\textbf{Craft:} blueprint (\emph{Wits}+Craft), brace a door (\emph{Body}+Craft)\\
\textbf{Melee:} break guard (\emph{Body}+Melee), bind a blade (\emph{Wits}+Melee)\\
\textbf{Ranged:} leading shot (\emph{Wits}+Ranged), loose in a squall (\emph{Spirit}+Ranged)\\
\textbf{Brawl:} grapple (\emph{Body}+Brawl), feint (\emph{Wits}+Brawl)\\
\textbf{Tactics:} ambush lane (\emph{Wits}+Tactics), coordinate a retreat (\emph{Presence}+Tactics)\\
\textbf{Diplomacy:} audience (\emph{Presence}+Diplomacy), draft terms (\emph{Wits}+Diplomacy)\\
\textbf{Sway:} haggle (\emph{Presence}+Sway), sell a risky plan (\emph{Presence}+Sway)\\
\textbf{Deception:} lie under scrutiny (\emph{Presence}+Deception), misdirect guard (\emph{Wits}+Deception)\\
\textbf{Performance:} captivate a crowd (\emph{Presence}+Performance), mimic voice (\emph{Wits}+Performance)\\
\textbf{Subterfuge:} talk past a checkpoint (\emph{Presence}+Subterfuge), case a routine (\emph{Wits}+Subterfuge)\\
\textbf{Streetwise:} find a fence (\emph{Presence}+Streetwise), sift rumors (\emph{Wits}+Streetwise)\\
\textbf{Arcana:} read a ward anchor (\emph{Wits}+Arcana), hold a rite steady (\emph{Spirit}+Arcana)\\
\textbf{Mechanics:} diagnose a lock (\emph{Wits}+Mechanics), field-rig a pump (\emph{Body}+Mechanics)\\
\textbf{Investigation:} reconstruct a scene (\emph{Wits}+Investigation), interview a witness (\emph{Presence}+Investigation)\\
\textbf{Lore:} cite local custom (\emph{Presence}+Lore), perform a rite (\emph{Spirit}+Lore)\\
\textbf{Nature:} read weather (\emph{Wits}+Nature), calm a mount (\emph{Presence}+Nature)\\
\textbf{Medicine:} stabilize a wound (\emph{Wits}+Medicine), cut away rot (\emph{Body}+Medicine)\\
\textbf{Command:} rally allies (\emph{Presence}+Command), hold the line (\emph{Spirit}+Command)
\end{multicols}
\vspace{0.5em}
\textbf{Fast Boundaries:}
\begin{itemize}
  \item \textbf{Locks \& Traps:} Mechanical = Mechanics + Attribute; Arcane = Arcana + Attribute.
  \item \textbf{People vs. Mechanisms:} Subterfuge for social infiltration; Stealth to avoid detection; Mechanics or Arcana to bypass a physical or magical barrier.
  \item \textbf{Formal vs. Informal:} Diplomacy for courts and treaties; Sway for bargaining and personal persuasion.
\end{itemize}
\end{tcolorbox}
