\clearpage
\section*{Appendix I: Session Zero Checklist Handout}
\addcontentsline{toc}{section}{Appendix I: Session Zero Checklist Handout}
\label{app:session-zero-handout}
\index{Session Zero!handout}

\begin{tcolorbox}[colback=green!5!white, colframe=green!75!black, title=Session Zero Checklist, fonttitle=\bfseries,breakable]
Use this checklist during your Session Zero. Write down the answers where everyone can see them.
\begin{enumerate}
  \item \textbf{Lines \& Veils} -- What content is off-limits (Lines) or only implied (Veils)?
  \item \textbf{Tone} -- Cinematic action? Grim survival? Political intrigue? Dark horror? Mix of several?
  \item \textbf{Campaign length} -- One-shot, short arc (3--6 sessions), or long campaign (12+ sessions)?
  \item \textbf{Character hooks} -- Each player shares one goal and one fear for their character.
  \item \textbf{Patron interest} -- Which Patrons (if any) are already watching the party?
  \item \textbf{Starting situation} -- Where does the story begin? (Use the Crown Spread or a simple opener.)
  \item \textbf{House rules} -- Any optional subsystems (Threat Pool, Deck of Consequences, etc.) you plan to use?
  \item \textbf{Scheduling \& attendance} -- How often do you meet? What happens if a player misses a session?
  \item \textbf{Player-GM contract} -- Remind the table: ``We are co-creators, not adversaries.''
  \item \textbf{Resource mindset} -- ``Boons are not XP tokens. Spend them. Embrace failure.''
  \item \textbf{Patron expectation} -- ``They are allies who demand payment -- embrace it.''
  \item \textbf{Flavor is free} -- Encourage descriptive narration for all actions, not just magic.
\end{enumerate}
\end{tcolorbox}
