\clearpage
\section*{Appendix E: Safety and Inclusivity}
\addcontentsline{toc}{section}{Appendix E: Safety and Inclusivity}
\label{app:safety}

\begin{tcolorbox}[title=X-Card System, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{safety tools!X-Card}
\begin{itemize}
  \item \textbf{X-Card:} ``I'm not comfortable -- change direction immediately.'' No explanation required.
  \item \textbf{O-Card:} ``I'm excited -- lean into this element.''
  \item \textbf{N-Card:} ``I need a break -- pause the scene.''
\end{itemize}
Keep cards visible. Model appropriate use as the GM.
\end{tcolorbox}

\begin{tcolorbox}[title=Script Change, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{safety tools!script change}
Ask at appropriate moments:
\begin{itemize}
  \item \textbf{Pause:} ``Are you okay with this content?''
  \item \textbf{Rewind:} ``Would you like to go back and change something?''
  \item \textbf{Fast Forward:} ``Would you like to skip ahead?''
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Inclusive Worldbuilding, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{inclusivity}
\begin{itemize}
  \item Research real cultures thoroughly when drawing inspiration.
  \item Avoid stereotypes; build characters with depth and agency.
  \item Provide character options that do not rely on real-world cultural markers.
  \item Use high-contrast, readable materials for accessibility.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Conflict Resolution, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{conflict resolution}
When conflicts arise at the table:
\begin{enumerate}
  \item Pause gameplay.
  \item Discuss privately if needed.
  \item Focus on understanding, not assigning fault.
  \item Find collaborative solutions.
  \item Document any agreements.
\end{enumerate}
Common solutions: the GM's ruling stands for the session; research the rule later; adjust future play.
\end{tcolorbox}
