\clearpage
\section*{Appendix D: Campaign Management}
\addcontentsline{toc}{section}{Appendix D: Campaign Management}
\label{app:campaign}

\begin{tcolorbox}[title=Campaign Timer Examples, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{campaign timers!examples}
\begin{itemize}
  \item \textbf{Mandate Advancement:} Public victories, crisis resolution, recognition by powerful factions (see \ref{subsec:mandate}).
  \item \textbf{Crisis Advancement:} Rival faction gains influence, asset neglect, betrayal, scandal (see \ref{subsec:crisis}).
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Follower and Asset Condition States, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{follower conditions}\index{asset conditions}
\begin{description}
  \item[Maintained] Full capability, no penalties.
  \item[Neglected] --1 die penalty; narrative wear and tear.
  \item[Compromised] Unavailable until repaired or recovered through downtime.
\end{description}
See \ref{sec:followers-assets-gm} for detailed rules.
\end{tcolorbox}

\begin{tcolorbox}[title=Faction Loyalty Scale (Optional), colback=white!97!gray, colframe=black!80!gray,breakable]
\index{factions!loyalty scale}
Track faction standing on a simple --3 to +3 scale.
\begin{description}
  \item[--3 Enemy] Actively works against the party.
  \item[0 Neutral] Indifferent.
  \item[+3 Ally] Will sacrifice for the party.
\end{description}
Adjust by \(\pm 1\) after major interactions (saving or betraying the faction).
\end{tcolorbox}

\begin{tcolorbox}[title=Campaign Pacing (Arc Structure), colback=white!97!gray, colframe=black!80!gray,breakable]
\index{campaign pacing}
\begin{enumerate}
  \item \textbf{Introduction} (1--2 sessions): Establish stakes, hook the players.
  \item \textbf{Development} (2--4 sessions): Complications multiply, alliances form.
  \item \textbf{Climax} (1--2 sessions): Major confrontation, critical choices.
  \item \textbf{Resolution} (1 session): Consequences land, new status quo established.
\end{enumerate}
\end{tcolorbox}
