\chapter{Quick Reference and Tables}
\label{ch:gm-reference}
\index{quick reference}

This chapter collects quick-reference tables, optional rules, and GM tools. Use them as needed at the table. For additional player-facing reference, see the \textit{Player's Guide} appendices.

\clearpage
\section*{Appendix A: Core Reference Tables}
\addcontentsline{toc}{section}{Appendix A: Core Reference Tables}
\label{app:core-tables}

\begin{longtable}{lll}
\caption{Outcome Matrix} \label{tab:outcome-matrix} \\
\toprule
\textbf{Case} & \textbf{Name} & \textbf{Guidance} \\
\midrule
\endfirsthead
\multicolumn{3}{c}%
{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Case} & \textbf{Name} & \textbf{Guidance} \\
\midrule
\endhead
\bottomrule
\endfoot
$S \geq DV$, $C = 0$ & Clean Success & Deliver the intent crisply. \\
$S \geq DV$, $C > 0$ & Success \& Cost & Grant intent; spend or bank SB for complications. \\
$0 < S < DV$ & Partial & Progress; award 1 Boon. \\
$S = 0$ & Miss & No progress; award 2 Boons; spend or bank SB. \\
\end{longtable}
\index{outcome matrix}\index{clean success}\index{success with sb}\index{partial success}\index{miss}

\begin{tcolorbox}[title=Story Beat (SB) Spend Menu, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{story beats!spending}
\begin{itemize}
    \item \textbf{1 SB}: Minor pressure -- noise, trace, \(+1\) segment on a relevant timer.
    \item \textbf{2 SB}: Moderate setback -- alarm raised, lose cover or advantage, lesser foe appears.
    \item \textbf{3 SB}: Serious trouble -- reinforcements arrive, key gear breaks temporarily, tick a campaign front.
    \item \textbf{4+ SB}: Major turn -- trap springs, authority arrives, scene shifts dramatically.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Position Descriptions, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{position!descriptions}
\begin{itemize}
    \item \textbf{Dominant}: Act on your terms. Consequences are manageable.
    \item \textbf{Controlled}: Normal risk and reward. Failure has a real cost.
    \item \textbf{Desperate}: Odds stacked against you. Failure could be catastrophic.
\end{itemize}
\end{tcolorbox}

\begin{longtable}{cll}
\caption{Difficulty Ladder} \label{tab:dv-ladder} \\
\toprule
\textbf{DV} & \textbf{Name} & \textbf{When to Use} \\
\midrule
\endfirsthead
\multicolumn{3}{c}%
{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{DV} & \textbf{Name} & \textbf{When to Use} \\
\midrule
\endhead
\bottomrule
\endfoot
2 & Routine & Clear intent, modest stakes, controlled environment. \\
3 & Pressured & Time pressure, mild opposition, partial information. \\
4 & Hard & Hostile conditions, active opposition, precise timing. \\
5+ & Extreme & Multiple constraints, high stakes, dramatic failure. \\
\end{longtable}
\index{difficulty value!ladder}

\begin{tcolorbox}[title=Boon Economy Quick Guide, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{boons!economy}
\begin{itemize}
  \item \textbf{Holding cap:} 5 Boons maximum.
  \item \textbf{Carryover:} At the end of each scene, reduce held Boons to a maximum of 2.
  \item \textbf{Conversion:} 2 Boons \(\rightarrow\) 1 XP, once per session during downtime (max 2 XP per session via conversion).
  \item \textbf{Spending:} Re-roll a die, activate an asset, improve Position, or power certain Rites.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Mechanical Constraints, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{assist limit}\index{follower actions}
\begin{itemize}
  \item \textbf{Assist Max:} \(+3\) dice total per roll (Exceptional Coordination allows \(+4\) from one follower).
  \item \textbf{Boon Max:} 5 total; 2\(\rightarrow\)1 XP conversion once per session (max 2 XP via conversion).
  \item \textbf{Follower Independent Action:} Once per scene (party-wide).
\end{itemize}
\end{tcolorbox}

\begin{longtable}{llll}
\caption{Harm Levels and Effects} \label{tab:harm} \\
\toprule
\textbf{Level} & \textbf{SB Generation} & \textbf{Penalty} & \textbf{Recovery} \\
\midrule
\endfirsthead
\multicolumn{4}{c}%
{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Level} & \textbf{SB Generation} & \textbf{Penalty} & \textbf{Recovery} \\
\midrule
\endhead
\bottomrule
\endfoot
Minor & 1 SB (next 2 rolls) & --1 die (related actions) & Rest or basic care \\
Moderate & 1 SB (next roll) & --1 die (most actions) & Medical treatment \\
Severe & 2 SB (next roll) & --2 dice (most actions) & Extended care \\
Critical & 3 SB (next roll) & Incapacitated & Major intervention \\
\end{longtable}
\index{harm!levels}

\begin{tcolorbox}[title=Range Bands, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{range bands}
\begin{description}
\item[Close] Arm's length, grappling distance.
\item[Near] Same room or immediate area.
\item[Far] Visible but not immediately reachable.
\item[Absent] Off-screen; requires a scene change.
\end{description}
\end{tcolorbox}

\begin{tcolorbox}[title=Movement Actions, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{movement actions}
\begin{itemize}
  \item \textbf{Move}: Shift one range band (Close \(\leftrightarrow\) Near or Near \(\leftrightarrow\) Far).
  \item \textbf{Dash}: Shift two bands as your full action.
  \item \textbf{Disengage}: Test to leave an enemy's Close range without provoking an opportunity attack.
  \item \textbf{Charge / Skirmish}: Full-turn actions that move two bands and attack, costing 1 Fatigue.
\end{itemize}
\end{tcolorbox}

\begin{tcolorbox}[title=Supply Timer States, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{supply timer}
\begin{description}
\item[Full (0 segments)] Well-provisioned, no penalty.
\item[Low (2 segments)] Minor friction -- resource checks at --1 die.
\item[Dangerously Low (3 segments)] Each PC gains Fatigue 1.
\item[Empty (4 segments)] Severe penalties -- starvation or thirst risk.
\end{description}
\end{tcolorbox}

\begin{tcolorbox}[title=Travel Timer Sizes, colback=white!97!gray, colframe=black!80!gray,breakable]
\index{travel timers}
\begin{description}
\item[4 segments] Short, straightforward leg.
\item[6 segments] Standard leg.
\item[8 segments] Extended or complex journey.
\item[10 segments] Epic or highly dangerous passage.
\end{description}
\end{tcolorbox}

\begin{longtable}{lll}
\caption{XP Costs} \label{tab:xp-costs} \\
\toprule
\textbf{Improvement} & \textbf{Cost} & \textbf{Downtime} \\
\midrule
\endfirsthead
\multicolumn{3}{c}%
{{\tablename\ \thetable{} -- continued}} \\
\toprule
\textbf{Improvement} & \textbf{Cost} & \textbf{Downtime} \\
\midrule
\endhead
\bottomrule
\endfoot
Attribute increase & new rating \(\times\) 3 XP & new rating days \\
Skill increase & new level \(\times\) 2 XP & new level days \\
On-Screen Follower & Cap\(^2\) XP & 1--3 days \\
Minor Asset & 4 XP & 1 day \\
Standard Asset & 8 XP & 1 week \\
Major Asset & 12 XP & 1 month \\
\end{longtable}
\index{XP!costs}

\begin{tcolorbox}[title=Upkeep (Condensed), colback=white!97!gray, colframe=black!80!gray,breakable]
\index{upkeep}
\textbf{Frequency:} Once per downtime period.
\begin{itemize}
  \item \textbf{Efficient (higher XP, less time):} Upkeep XP = \(\max(1, \frac{\text{Acquisition XP}}{3})\).
  \item \textbf{Intensive (1 XP, dedicated action):} One full downtime action.
  \item \textbf{Failure to pay:} Follower becomes Wary (or Seized if already Wary); Asset becomes Neglected (or Compromised if already Neglected).
\end{itemize}
\end{tcolorbox}
