\section{Inaea — Angel of the Spider}

\subsection{Domain and Epithets}

Inaea sits in patient silence at the centre of the unseen lattice. Where others see isolation, she sees threads—debt, affection, rivalry, loyalty—all binding mortals to one another. She is comforter and conspirator alike: the hearth-mother who eases grief with a warm hand, and the shadow-spinner who knots fates so they cannot slip free. Those who serve her learn to pull strands, tying allies closer and ensnaring foes until mercy or trap is inevitable.

The spider does not rush. It weaves, waits, and feels the tremor of every caught fly. So too does Inaea. Her mercy is never without cost; her kindness is a thread that, once accepted, can be pulled taut at a moment of her follower's choosing.

\subsection{Cultural Origins}

Inaea is honored wherever relationships are woven and bonds are binding:

\begin{itemize}
\item In \textbf{Aelaerem}, she is the Kettle-Spinner. Grandmothers invoke her to keep families together, tying red thread around the wrists of newborns.
\item In \textbf{Lethai} courts, she is the Silk Vigil, invoked before marriages and treaties. Her symbol is a single strand of unbroken silk.
\item In \textbf{Silkstrand}, she is the Dyed Web. Dyers' guilds leave a single strand of crimson silk outside their windows on the first of spring.
\item In \textbf{Vhasia}, she is the Tether-Mother, invoked by those who have been separated from loved ones by war. Her shrines are bridges and border stones.
\end{itemize}

\subsection{Warrior Archetype: The Web-Warrior}

A Runic Warrior of Inaea is not a front-line combatant in the traditional sense. They are a \textit{Web-Warrior}—a fighter who binds foes in invisible threads, who controls the battlefield through entanglement and restraint, who never needs to draw steel because the enemy is already caught. Their Codex is a spider-silk cord or a loom-shuttle. Their Thiasos is a spider-spirit that weaves, a cord that tightens when lies are spoken, or a sentient blade wrapped in silk that never dulls.

\subsection{Codex Form}

The Web-Warrior's Codex is a tool of binding:

\begin{itemize}
\item \textbf{Spider-Silk Cord:} A length of pale, almost invisible silk. It is stronger than steel. It tightens when a promise is broken.
\item \textbf{Loom-Shuttle:} A wooden shuttle carved from hawthorn. It weaves threads that are not quite physical—bonds of obligation, memory, and fate.
\item \textbf{Threaded Needle:} A silver needle with a red thread already threaded. It can stitch wounds—or stitch lips shut.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Spider-Spirit:} A translucent spider the size of a hand. It weaves threads between people, binding them in webs of obligation. It can sense hidden relationships (debt, affection, rivalry) within Near range.
\item \textbf{Binding Cord:} A length of grey cord that tightens when a promise is broken. It can be tied around a wrist to witness an oath.
\item \textbf{The Loom-Shuttle (Sentient):} A shuttle that speaks in the click of weaving. It can tie a Thread of Mercy between two parties—they suffer -1 die if they harm each other.
\end{itemize}

\subsection{Core Mechanic: The Web of Obligation}

When the Web-Warrior names a relationship between two NPCs (friend, rival, debtor, lover), the GM must reveal one secret about that relationship—a hidden grievance, a secret loyalty, an unpaid debt. This can be used once per scene.

When the Web-Warrior ties a physical thread (cord, ribbon, hair) between two objects or people, they may spend 1 Fatigue to create a \textit{Binding}. The binding lasts until the thread is cut. While bound, the two parties suffer -1 die if they act against each other's interests.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Hearth-Thread Knot (4 XP)}

\textbf{Tags:} [BOND], [PROTECT], [HEAL]

\textbf{Threshold:} A thread tied around a wrist or a shared cup.

Bind yourself and one willing target with a sympathetic thread. Both gain +1 die to Aid each other this scene. If either takes Harm, the other may reduce it by 1 (to a minimum of 0) by suffering that same Harm instead.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Snaring Filament (5 XP)}

\textbf{Tags:} [TRAP], [AREA], [BIND]

\textbf{Threshold:} A knotted cord or woven scrap.

Lay an invisible snare in a lane, doorway, or narrow passage. The first hostile creature to cross suffers -2 dice on their next action and must test Resolve (DV 2) or lose their movement for that turn (tangled).

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Strand of Inevitability (8 XP)}

\textbf{Tags:} [BOND], [FATE], [MANIPULATION]

\textbf{Threshold:} Three coloured threads (white, black, red) braided together.

Link two actors with a visible or invisible strand. For the rest of the scene, whenever one acts, the other is dragged into consequence. You may trigger this linkage once to either force 1 SB (Hearts) on the linked target after they perform a significant action, or gain +2 dice on a Setup, Predict, or Manipulate roll that exploits the link.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Spinner's Debt (9 XP)}

\textbf{Tags:} [VOW], [DEBT], [BOND], [COMMAND]

\textbf{Threshold:} A silver coin and a strand of your own hair, knotted together.

Offer a willing creature a boon: they may immediately clear 1 Fatigue or gain +1 die on their next roll. In return, they accept a Promise Clock [4] named "Debt to the Web." You may advance this clock by 1 segment at any time as a free action (once per scene). When the clock fills, you may demand a single service from them—not impossible, but significant.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Merciful Severing (13 XP)}

\textbf{Tags:} [HEAL], [DISPEL], [MERCY], [DEBT]

\textbf{Threshold:} A white thread burned to ash, spoken with a name.

Sever one harmful tie—a curse, a toxic obsession, a coercive bond, or a supernatural debt. Remove one Condition or 1 SB tied to it. This is an act of genuine mercy; the target may be unwilling, but the thread can still be cut (Resolve test DV 4 to resist).

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Tug the Strand (4 XP)}

\textbf{Requirement:} Inaea

Once per session, name a relationship between two NPCs (friend, rival, debtor, lover). The GM must reveal one secret about that relationship—a hidden grievance, a secret loyalty, an unpaid debt.

\subsubsection{Binding Presence (6 XP)}

\textbf{Requirement:} Inaea, Presence 3+

Once per session, you may compel a promise from a target unless they pass a Resolve test (DV 4). The promise is not magical—but breaking it will be noticed.

\subsubsection{Mercy's Thread (8 XP)}

\textbf{Requirement:} Inaea, Tier II+

When you show mercy to an enemy (sparing their life, binding their wounds, offering sanctuary), you may tie a thread between you. The target gains a Promise Clock [4] named "Mercy's Debt." When the clock fills, they must perform one act of kindness for someone else—not for you. The spiral continues.

\subsection{Roleplaying a Web-Warrior}

\begin{itemize}
\item You do not fight fair. You bind, entangle, and restrain. The enemy does not need to die—only to be unable to act.
\item You may be called to mediate disputes, to witness oaths, or to untangle relationships that have become knotted and harmful.
\item You may struggle with the weight of obligation. Every thread you tie is a leash. Every mercy you offer is a debt. The web does not distinguish between kindness and control.
\item You may be asked to cut a thread—to end a relationship, to sever a bond, to let someone go free. The spider does not cut. The spider waits.
\end{itemize}

\subsection{Sample NPC: Weaver Anara of the Silk Vigil}

\textbf{Archetype:} Web-Warrior

\textbf{Patron:} Inaea

\textbf{Codex:} A spider-silk cord tied around her wrist

\textbf{Thiasos:} A spider-spirit that weaves threads between people

\textbf{Personality:} Patient, watchful, quietly amused. She has never lost a negotiation. She has never needed to fight.

\textbf{Conflict:} A thread she tied twenty years ago—between a mother and a daughter—has tightened. The daughter is dying. The mother is far away. The thread remembers. Anara must decide: cut it, or let it hold until the end.

\subsection{GM Guidance: Intrusions for Inaea}

\begin{itemize}
\item A spider lowers itself onto the Web-Warrior's shoulder. In its web is a single perfect word—the name of someone who needs their help.
\item Every handshake the Web-Warrior offers for the next day feels like grasping a thread. They sense whether the person is lying—but the person also senses their scrutiny.
\item A knot the Web-Warrior tied earlier is now impossible to undo without cutting. Cutting it severs a minor bond—they lose 1 Boon.
\item The Web-Warrior dreams of a web stretching across the sky. In the centre, a spider watches. It does not move. It does not need to.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
