\section{Isoka — Angel of Serpents}

\subsection{Domain and Epithets}

Isoka, sister to Ikasha (Shadow) and Inaea (Mercy), completes the Triad of Transformation. Where thresholds and compassion mark her sisters' domains, Isoka teaches that every self is temporary—identity is a skin to be shed so growth can continue. Her serpents are omens and teachers: each cast skin a lesson in release; each venom a catalyst for necessary change. Those who walk her path become alchemists of the self, embracing dissolution as the first motion of rebirth.

Isoka does not judge. She offers the knife and the poison both. The choice is the seeker's.

\subsection{Cultural Origins}

Isoka is honored wherever transformation is embraced or inflicted:

\begin{itemize}
\item In \textbf{Ashaan}, she is the Uncoiled Chain, invoked by those who seek to shed their past—slaves, fugitives, the desperate. Her shrines are hidden in the cracks of the Veil.
\item In \textbf{Lethai} courts, she is the Shed Vigil, invoked before trials of metamorphosis. Her symbol is a serpent eating its own tail.
\item In \textbf{Sekogo}, she is the Venom-Mother, whose priests use snake venom in rites of initiation. To survive is to be reborn.
\item In \textbf{Zakov}, she is the Tide-Shedder. Pirates leave a lock of hair and a name at her sunken temple; eels eat both. The name is forgotten, and the pirate gains a year of incredible luck—but the eels remember.
\end{itemize}

\subsection{Warrior Archetype: The Serpent-Blade}

A Runic Warrior of Isoka is not a weaver of webs or a keeper of secrets. They are a \textit{Serpent-Blade}—a warrior who sheds identity like skin, who transforms in the midst of battle, who uses venom as a weapon and change as a shield. Their Codex is a shed snakeskin or a fang. Their Thiasos is a serpent-spirit that whispers of change, a venomous familiar, or a sentient blade that drips with metamorphic poison.

\subsection{Codex Form}

The Serpent-Blade's Codex is a testament to transformation:

\begin{itemize}
\item \textbf{Shed Snakeskin:} A complete snakeskin, preserved and carried in a pouch. It is delicate. It is a record of a self that no longer exists.
\item \textbf{Fang:} A hollow serpent's fang on a leather cord. It can deliver venom—or draw it out.
\item \textbf{Scale:} A single iridescent scale from a serpent that shed its skin in a threshold (doorway, crossroads, grave). It is warm when change is near.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Serpent-Spirit:} A translucent snake that coils around the Runekeeper's arm. It can sense when someone is about to change—to leave, to betray, to die.
\item \textbf{Venomous Familiar:} A small snake, spider, or scorpion. Its venom can heal or harm. The Runekeeper is immune.
\item \textbf{The Shed Fang (Sentient):} A hollow fang that speaks in a sibilant whisper. It can deliver a dose of venom (Harm 1 + Poisoned condition) once per session.
\end{itemize}

\subsection{Core Mechanic: The Shedding}

When the Serpent-Blade successfully resists or breaks free from a Condition (Fear, Poison, Bound, etc.), they gain 1 Boon. The shedding is not escape—it is transformation.

When the Serpent-Blade helps another creature shed a Condition (through healing, comfort, or ritual), they may clear 1 Fatigue. The spiral of change continues.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Loosening Skin (4 XP)}

\textbf{Tags:} [HEAL], [ESCAPE], [CLEANSE]

\textbf{Threshold:} A discarded snakeskin or a strand of your own hair.

Shrug off a temporary burden. Remove one Condition (Fear, Poison, Fatigue, or a minor social stain) that you have acquired this scene. Alternatively, reroll one failed roll to escape restraint or pursuit.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Forked Tongue (5 XP)}

\textbf{Tags:} [DECEPTION], [BOOST], [RISK]

\textbf{Threshold:} A harmless lie whispered to a mirror.

For one scene, you may speak ambiguously without penalty. When you use Deception or Sway to conceal your true intent or pass as someone else, you may treat a Partial Success as a Clean Success (the lie holds). On a Miss, however, the GM gains 2 SB as your deceit unravels in an embarrassing way.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Shedded Self (8 XP)}

\textbf{Tags:} [TRANSFORM], [ERASURE], [ESCAPE], [DEBT]

\textbf{Threshold:} A complete change of clothing and a token from your former role.

You shed one persistent identity—a title, a bond, a debt, or a reputation. For the remainder of the scene, you are treated as a stranger to those who knew you in that role. You gain +2 dice on any roll to avoid detection, escape pursuit, or convince others that you are not yourself. Create a 6-segment Shed Identity clock.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of Venom's Paradox (9 XP)}

\textbf{Tags:} [POISON], [HEAL], [CURSE], [CLEANSE]

\textbf{Threshold:} A drop of venom from a serpent that has shed its skin within the last day.

Inflict or cure through poison's paradox. Choose one: cursed venom (touch a creature and inject venom; they suffer +1 Harm and must test Resolve (DV 4) or gain the Poisoned Condition) or mercy's antidote (touch a poisoned or diseased creature; the venom neutralizes the malady; remove one Poisoned or Diseased Condition, and the target clears 1 Fatigue).

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of Unremembered Metamorphosis (13 XP)}

\textbf{Tags:} [TRANSFORM], [ERASURE], [FATE], [BURDEN]

\textbf{Threshold:} A full identity kit, a serpent's shed skin, and a personal item from the life you are leaving.

Choose one: complete metamorphosis (after an hour of ritual, you emerge with a new face, voice, and documented identity; for one week, immune to recognition by anyone who knew your previous self) or cast-off history (after a full night's ritual burning all mundane records of your old life, your previous identity becomes unreliable; common records and casual memories become vague).

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Shed Skin (4 XP)}

\textbf{Requirement:} Isoka

Once per session when you are Bound, Grappled, or suffering from a Condition that restricts movement, you may spend an action and mark 1 Fatigue to shed that restriction. Describe how you slip free like a snake shedding its skin.

\subsubsection{Venomous Strike (6 XP)}

\textbf{Requirement:} Isoka

When you successfully strike a target with a melee attack, you may choose to deliver a dose of venom. The target suffers +1 Harm and must test Resolve (DV 4) or gain the Poisoned Condition. You may use this once per scene.

\subsubsection{Momentum of the Molt (8 XP)}

\textbf{Requirement:} Isoka, Body 3+

When you successfully resist or break free from a Condition, you gain a burst of furious energy. Your next action within the same scene gains +1 die and treats its Position as one step higher (Desperate → Controlled, Controlled → Dominant). You may use this once per scene.

\subsection{Roleplaying a Serpent-Blade}

\begin{itemize}
\item You do not cling to who you were. That self is a shed skin. The new self is still forming.
\item You may be called to help others change—to leave abusive relationships, to shed destructive habits, to become someone new. The knife is sharp. The venom burns. Transformation is not gentle.
\item You may struggle with the loss of continuity. If you shed your identity too often, you may forget who you were at the start. Isoka does not promise memory. She promises change.
\item You may be asked to stop—to stay one self, to keep a name, to remember a promise. The serpent does not cling. The serpent sheds.
\end{itemize}

\subsection{Sample NPC: Copperhead of the Shed Vigil}

\textbf{Archetype:} Serpent-Blade

\textbf{Patron:} Isoka

\textbf{Codex:} A shed snakeskin, preserved in a leather pouch

\textbf{Thiasos:} A serpent-spirit that whispers when change is near

\textbf{Personality:} Soft-spoken, watchful, patient. She has shed her name seven times. She is not sure which one was real.

\textbf{Conflict:} A former self she thought she had shed has returned—not as a memory, but as a person. Her old name walks the roads. Her old debts are calling. The serpent says: shed again. The heart says: remember.

\subsection{GM Guidance: Intrusions for Isoka}

\begin{itemize}
\item A piece of the Serpent-Blade's old self returns—a letter from a dead lover, a scar they had forgotten, a name spoken in a dream.
\item The Serpent-Blade's shadow moves independently, pointing at people who are about to change—to leave, to betray, to die.
\item The Serpent-Blade begins to shed skin overnight. If they burn the cast, they forget one minor memory. If they keep it, someone will steal it.
\item The Serpent-Blade dreams of a serpent eating its own tail. The circle is endless. The change is never complete.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
