\section{Malachai — The Cruel Messenger}

\subsection{Domain and Epithets}

Malachai was once a divine messenger—a herald of painful truths, sent to deliver news that could not be softened. She was not cruel; she was necessary. But necessity wears. In time, she began to soften the truth, offering false hope alongside real curses. For this, the gods bound her in chains of black iron beneath the cursed city of Zahk'Kozahn.

Now she waits. Chained but not contained. Imprisoned but not silent. Her voice reaches across the world, offering power to the weak, hope to the hopeless, and sweet oblivion to those who cannot bear to face the dawn. Vampirism, lycanthropy, and other cursed conditions flow from her influence. Her followers learn that power demands suffering, mastering gifts that devour from within.

\subsection{Cultural Origins}

Malachai appears wherever desperation meets opportunity:

\begin{itemize}
\item In \textbf{Linn}, she is the Salt-Fang. Shipwreck survivors sometimes find a silver coin in their mouth. Those who spend it gain sea-lungs (breathe underwater) but develop a hunger for raw flesh.
\item In \textbf{Valewood}, she is the Thorn-Giver. Hunters who curse a rival may find their arrows always strike true—but the blood on the arrowhead never dries.
\item In \textbf{Ecktoria}, she is the Ur-Heretic, the source of all corruptions. Her followers are burned without trial.
\item In \textbf{Zakov}, she is the Tide-Shedder. Pirates leave a lock of hair and a written name at a sunken temple. Eels eat both. The name is forgotten, and the pirate gains a year of incredible luck—but the eels remember.
\end{itemize}

\subsection{Warrior Archetype: The Thorn-Sworn}

A Runic Warrior of Malachai is not a healer or a guardian. They are a \textit{Thorn-Sworn}—a cursed warrior who has accepted power at a price, who carries their affliction like a second skin, who wields the very thing that destroys them. Their Codex is a silver coin that never spends or a thorn that draws blood. Their Thiasos is a Collector (a hollowed child with silver-wire eyes), a cursed hound, or a sentient blade that whispers corrupt bargains.

\subsection{Codex Form}

The Thorn-Sworn's Codex is a testament to cursed bargains:

\begin{itemize}
\item \textbf{Silver Coin:} A silver coin that is always warm. It can be spent once to reroll a failed roll—but the coin returns at dawn, heavier. Each use adds a scar.
\item \textbf{Thorn:} A black thorn that draws blood when touched. It never dries. It remembers every bargain struck.
\item \textbf{Cracked Symbol:} A broken symbol of a forgotten patron. It weeps black ichor when a curse is invoked.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Collector:} A hollowed child with milk-white eyes and silver wire sewing the lids shut. It does not speak. It watches. It collects debts when they come due.
\item \textbf{Cursed Hound:} A black hound with eyes like burning coals. It can track any creature that has broken an oath. It is never fully loyal—it serves the curse, not the curser.
\item \textbf{The Silver Coin (Sentient):} A coin that speaks in the voice of the last person who spent it. It can be offered as a bribe that cannot be refused—but the giver forgets what they paid for.
\end{itemize}

\subsection{Core Mechanic: The Cursed Bargain}

When the Thorn-Sworn accepts a curse (from Malachai, from a Rite, or from circumstance), they may mark 1 Fatigue to gain +1 die on their next action. The curse feeds the power.

When the Thorn-Sworn inflicts a curse on another, they may clear 1 Fatigue. The curse is transferred—but not forgotten.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Honeyed Binding (5 XP)}

\textbf{Tags:} [CURSE], [BOOST], [DEBT], [FEAR]

\textbf{Threshold:} Blood mixed with sweet wine, or a broken chain link.

Choose one: honeyed curse (target gains +2 dice to their next roll—counts as Triumph; they take 1 Corruption, unresistable) or binding curse (target gains +1 die to physical actions but suffers Fatigue 1 unless they perform a dominance act).

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the False Dawn (6 XP)}

\textbf{Tags:} [CURSE], [DEBT], [FATE]

\textbf{Threshold:} Meteorite iron shard blessed by false light.

Bestow a gift with a hidden curse. Target gains a significant advantage (e.g., +1 die to all rolls for one scene) but accrues a 4-segment Corruption clock. Each time they use the gift, tick the clock. When full, they suffer a related curse (GM chooses: festering wounds, beastly urges, or the attention of a Collector).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Cruel Transformation (9 XP)}

\textbf{Tags:} [CURSE], [TRANSFORM], [DEBT], [FEAR]

\textbf{Threshold:} Holy water mixed with graveyard dust.

Infect a target with a supernatural condition (vampiric hunger, bestial rage, shadow-bind). Lasts one scene. Start a 6-segment Transformation clock. Each time the target resists feeding or transformation, tick the clock. When full, the condition becomes permanent.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Messenger's Burden (13 XP)}

\textbf{Tags:} [CURSE], [BOOST], [PRESSURE], [DEBT]

\textbf{Threshold:} Manacles worn while speaking her name.

Channel Malachai directly. Gain +2 dice to one action per beat this scene, but each use marks 1 SB as her influence spreads. Start an 8-segment Corruption's Grip clock.

\textbf{Obligation:} 3

\subsubsection{High Rite: Rite of the Chained Angel's Tribute (14 XP)}

\textbf{Tags:} [CURSE], [AREA], [SACRIFICE], [FATE]

\textbf{Threshold:} A chapel bell cracked by divine lightning, or a silver coin that cannot be spent.

Corrupt a sacred space or consecrate a site to Malachai. Within the zone: supernatural effects (curses, dark magic) gain +1 Effect; blessings and healing spells generate +1 SB on a failure; any creature that accepts a gift from Malachai within the zone immediately ticks a Corruption clock twice.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Borrowed Scars (6 XP)}

\textbf{Requirement:} Malachai

Once per session, transfer a Condition or 1 Fatigue from an ally to yourself. In exchange, gain +1 die on your next roll against the source of that affliction. Mark 1 Obligation after using this talent.

\subsubsection{Thorn's Memory (4 XP)}

\textbf{Requirement:} Malachai

You remember every curse you have inflicted. You can recall the face, the name, and the exact terms of any bargain you have struck with Malachai. This is a passive ability—and a burden.

\subsubsection{Cursed Resilience (8 XP)}

\textbf{Requirement:} Malachai, Body 3+

When you are under the effect of a curse (including your own), you gain +1 Armor. The curse protects its own—so that it may continue to collect.

\subsection{Roleplaying a Thorn-Sworn}

\begin{itemize}
\item You do not make promises lightly. Every promise is a curse waiting to bloom.
\item You may be called to deliver painful truths, to collect debts, or to offer desperate hope to the hopeless. The coin is always warm. The thorn never dries.
\item You may struggle with the hunger. The curse feeds. The more you use it, the more it wants. Malachai does not warn. She offers.
\item You may be asked to stop—to break a curse, to refuse a bargain, to let someone go free. The chains are patient. They do not release.
\end{itemize}

\subsection{Sample NPC: The Silver-Tongued of the Drowned Quarter}

\textbf{Archetype:} Thorn-Sworn

\textbf{Patron:} Malachai

\textbf{Codex:} A silver coin that is always warm

\textbf{Thiasos:} A Collector with milk-white eyes that never blink

\textbf{Personality:} Charming, desperate, self-destructive. He has made nine bargains with Malachai. He does not remember the terms of the first.

\textbf{Conflict:} A curse he inflicted years ago has returned—not on the target, but on the target's child. The child is innocent. The curse does not care. The coin is warm. The Collector watches.

\subsection{GM Guidance: Intrusions for Malachai}

\begin{itemize}
\item A gift the Thorn-Sworn gave earlier curdles. The recipient now suffers a visible curse—a limp, a stutter, a shadow that moves without them.
\item The moon turns red for one night. All lycanthropes within Near must test Resolve (DV 5) or lose control.
\item A messenger arrives with a letter sealed in black wax. It is from someone the Thorn-Sworn cursed years ago. They are coming.
\item The Thorn-Sworn's silver coin is warm—warmer than usual. Malachai is paying attention. A Collector is on its way.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
