\section{Varnek Karn — The Death's Negotiator}

\subsection{Domain and Epithets}

Varnek Karn dwells in the grey between life and death, where the newly departed linger with unfinished business and elder shades guard secrets dearer than gold. He is the broker who speaks across the veil, the mediator of disputes between living and dead, and the archivist of every promise sworn beyond the grave. His servants practice post-mortem diplomacy: they negotiate, consult, and contract with spirits as equals—mindful that the dead are patient and rarely generous.

Yet Varnek Karn is also the patron of those who \textit{raise} the dead, not through force but through bargain. The dead may be persuaded to walk again, to fight for a season, or to guard a threshold, but every such compact is a debt. The dead demand their due: a memory, a future, a name, or simply the promise that the living will remember them.

\subsection{Cultural Origins}

Varnek Karn is invoked wherever the living must treat with the dead:

\begin{itemize}
\item In \textbf{Aeler}, he is the Breath-Accountant, who audits the spiritual ledgers of the Deep Vaults and ensures that the dead are not forgotten.
\item In \textbf{Ykrul}, he is the Bone-Speaker, whose drum calls ancestors to council before any major migration.
\item In \textbf{Vhasia}, he is the Silent Notary, who witnesses death-bed testaments and ensures their terms are honoured.
\item In \textbf{Thepyrgos}, he is the Archive Keeper, whose shelves hold the names of those who died before they could pay their debts.
\end{itemize}

\subsection{Warrior Archetype: The Cairn-Knight}

A Runic Warrior of Varnek Karn is not a necromancer who commands unwilling spirits. They are a \textit{Cairn-Knight}—a death-negotiator who fights alongside ancestral shades, who carries the weight of the dead into battle, who pays in blood what others pay in gold. Their Codex is a rune-carved axe or a burial stone. Their Thiasos is an ancestral shade, a cairn-wight, or a sentient blade forged from the iron of a funeral pyre.

\subsection{Codex Form}

The Cairn-Knight's Codex is a testament to ancestral memory:

\begin{itemize}
\item \textbf{Rune-Carved Axe:} A battle axe whose haft is carved with the names of the Runekeeper's ancestors. Each name is a Rite waiting to be invoked.
\item \textbf{Burial Stone:} A flat stone taken from a cairn. It is cold to the touch. It warms when a debt to the dead is near.
\item \textbf{Wergild Bracelet:} A bronze bracelet with notches for each blood-debt paid. Each notch is a life avenged.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Ancestral Shade:} A translucent figure in ancestral garb—armor, robes, or furs. It fights beside the Runekeeper. It can pass through walls. It cannot be harmed by mundane weapons.
\item \textbf{Cairn-Wight:} A small spirit bound to a burial stone. It can sense when a corpse is disturbed. It can whisper the names of the dead.
\item \textbf{The Grave-Token (Sentient):} A polished skull the size of a fist. It speaks in the voice of the Runekeeper's own dead. It knows the name of every ancestor within three generations.
\end{itemize}

\subsection{Core Mechanic: The Ancestral Debt}

When the Cairn-Knight invokes a Rite that calls upon the dead, they may mark 1 Obligation to gain +1 die. The ancestors remember—and they expect repayment.

When the Cairn-Knight settles a blood-debt (wergild, revenge, or a fulfilled death-bed promise), they may clear 1 Obligation. The dead are patient, but they are not without mercy.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Crossing Fee (4 XP)}

\textbf{Tags:} [SPIRIT], [KNOWLEDGE]

\textbf{Threshold:} A coin for the tongue of the deceased, or a lock of hair from the bereaved.

Open speech with a specific dead individual. Gain +1 die to Investigation or Lore when seeking knowledge they once held.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Ancestral Consultation (5 XP)}

\textbf{Tags:} [SPIRIT], [TRUTH], [FATE]

\textbf{Threshold:} A personal token of the departed, or their favoured food or drink.

Ask one question of an ancestral spirit regarding family history, hidden legacies, or inherited debts.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Mediated Pact (8 XP)}

\textbf{Tags:} [VOW], [BOND], [COMMAND], [SPIRIT]

\textbf{Threshold:} Two copies of a written accord (one buried with the dead, one kept by the living).

Choose one: death-bound pact (bind a compact between a living signatory and a willing spirit; both gain +1 die when fulfilling the terms) or spirit medium (serve as a conduit for several spirits to speak; gain +2 dice to mediate disputes between living and dead).

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Unquiet Host (10 XP)}

\textbf{Tags:} [SUMMON], [AREA], [BOND], [DEBT]

\textbf{Threshold:} A battlefield relic (a broken sword, a skull, a scrap of uniform) and a handful of grave-earth.

Negotiate with the lingering dead of a place. Summon a Temporary Host—a spectral warband treated as a single Cap 3 Follower that grants +3 dice to a single combat or intimidation roll, or provides a free Setup action. When the Host departs, mark +2 Obligation or perform a Memorial Rite.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Final Testament (13 XP)}

\textbf{Tags:} [VOW], [FATE], [COMMAND], [DEATH]

\textbf{Threshold:} A dying person's last words recorded in their blood, witnessed by you.

Seal the last request with weight beyond life. Targets must pass Resolve (DV 4) to ignore it.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{The Silent Ledger (6 XP)}

\textbf{Requirement:} Varnek Karn

Once per session, touch a physical trace of a death (a body, a grave, an object that witnessed a passing) and instantly know one unresolved obligation connected to that death—a promise unfulfilled, a secret unspoken, a debt unpaid. Mark 1 Fatigue after using this talent.

\subsubsection{Final Witness (4 XP)}

\textbf{Requirement:} Varnek Karn, Spirit 3+

Once per session, when a sentient creature dies within Near range, you may ask the GM one question about that creature's final thoughts or unfulfilled wishes. The GM must answer truthfully.

\subsubsection{Reoccupied Vessel (10 XP)}

\textbf{Requirement:} Varnek Karn, Summoner's Path or Runekeeper's Path, Tier II+

You may perform a ten-minute ritual at the side of a corpse no more than three days dead. Negotiate with the spirit (which must be willing or sufficiently indebted). If it agrees, it reoccupies its body, becoming a Bound Vessel (Cap 3, Tier II-III) that serves for a set term or until a specific task is completed. Mark +2 Obligation.

\subsection{Roleplaying a Cairn-Knight}

\begin{itemize}
\item You do not command the dead. You negotiate with them. The dead are not slaves—they are creditors.
\item You may be called to settle inheritance disputes, to carry messages across the veil, or to negotiate with ghosts who refuse to leave.
\item You may struggle with the weight of ancestral memory. The dead have long memories. They remember every slight. They expect you to remember too.
\item You may be asked to let a spirit go—to release a debt, to forgive a wrong, to allow the dead to pass on. The ledger does not close easily.
\end{itemize}

\subsection{Sample NPC: Cairn-Keeper Harkan of the Bone-Speakers}

\textbf{Archetype:} Cairn-Knight

\textbf{Patron:} Varnek Karn

\textbf{Codex:} A rune-carved axe bearing the names of his ancestors

\textbf{Thiasos:} An ancestral shade that fought in the same wars he did

\textbf{Personality:} Quiet, watchful, burdened. He has carried the debts of his bloodline for forty years. He is not sure which debts are his.

\textbf{Conflict:} An ancestor's unpaid wergild has come due—not in gold, but in blood. The descendant of the victim demands satisfaction. The ancestor is long dead. The debt is not.

\subsection{GM Guidance: Intrusions for Varnek Karn}

\begin{itemize}
\item A ghost the Cairn-Knight once bargained with appears at an inopportune moment, demanding that they witness a new death or settle an old account.
\item A corpse the Cairn-Knight handled begins to whisper in their sleep, reciting the terms of a contract they thought was closed.
\item A funeral the Cairn-Knight attended is suddenly disputed—someone claims the deceased was not of sound mind when they made their final testament.
\item The Cairn-Knight's grave-token is warm. An ancestor is trying to speak. The message is urgent. The message is not kind.
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
