\section{The High Mother — The Unbroken Cord}

\subsection{Domain and Epithets}

The High Mother is not a patron in the traditional sense. She is the centre of the Cord—a vast, non-physical tangle of grey thread that records every secret whispered, every grief accepted, every name surrendered. She sits in the House Without Doors, neither alive nor dead, sustained by the accumulated debt of every Daughter who has ever served. The Daughters of the Unbroken Cord are her hands, her eyes, her witnesses. They appear as midwives, hospice nurses, and deathbed attendants. Their mercy has a ledger. Their comfort has a price.

The High Mother does not speak. The Cord weaves.

\subsection{Cultural Origins}

The High Mother is invoked wherever the Cord has root:

\begin{itemize}
\item In \textbf{Aelaerem}, she is the Hearth-Mother, the keeper of red threads tied around newborns' wrists. Her Daughters are midwives and hedge-witches.
\item In \textbf{Vhasia}, she is the Silent Notary, invoked by those who need a witness that will never speak. Her Daughters tend the dying and record last words.
\item In \textbf{Sidhi} ports, she is the Cord-Keeper, whose Daughters run hospices that ask no payment—but record every name in a ledger that never dries.
\item In \textbf{Ecktoria}, she is the Unbound Chain, feared by the Inquisition. Her Daughters are hunted—but never caught.
\end{itemize}

\subsection{Warrior Archetype: The Bloodsworn}

A Runic Warrior of the High Mother is not a negotiator of ancestral debts or a keeper of ancestral shades. They are a \textit{Bloodsworn}—a Daughter of the Cord who has taken the covenant into battle, who stores Fatigue as knots in her flesh, who pays in blood what others pay in gold. Their Codex is a set of flesh-tattoos, each Rite woven into their skin. Their Thiasos is a Hollowed child or heartless bird—a silent witness that never blinks.

\subsection{Codex Form}

The Bloodsworn's Codex is a testament to fleshly covenant:

\begin{itemize}
\item \textbf{Flesh-Tattoos:} Rites tattooed onto the skin using ink mixed with ash from former Vessels and blood-milk from the Cord. The tattoos move—slowly, subtly—as debts accrue.
\item \textbf{Grey Cord:} A length of undyed wool tied around the wrist. Each knot is a promise kept. The cord tightens when the Bloodsworn is tempted to break faith.
\item \textbf{Scar-Plate:} A set of ritual scars across the back or chest. Each scar is a Rite invoked. The scars never fully heal.
\end{itemize}

\subsection{Thiasos Form}

\begin{itemize}
\item \textbf{Hollowed Child:} A child taken from a home that no longer remembers them, hollowed out by the High Mother's needle. Heart replaced with a bundle of grey thread. Eyes sewn open with silver wire. It witnesses. It never blinks.
\item \textbf{Heartless Bird:} A small grey bird with milk-white eyes. It cannot sing. It watches. It weeps blood when a debt comes due.
\item \textbf{The Scarlet Needle (Sentient):} A silver needle with a red thread already threaded. It speaks in whispers. It can stitch wounds closed—or stitch lips shut.
\end{itemize}

\subsection{Core Mechanic: The Cord's Witness}

When the Bloodsworn invokes a Rite, their Hollowed or Heartless Bird counts as a witness (improving Position or reducing DV). The Cord sees. The Cord remembers.

When the Bloodsworn marks Obligation, they may store 1 point of it as a visible knot in their flesh (a fresh scar, a weeping tattoo, a tightened cord). They may have up to Body rating knots at once. While carrying knots, they appear uninjured—but the weight is real.

\subsection{Sample Rites}

\subsubsection{Low Rite: Rite of the Unspoken Grief (4 XP)}

\textbf{Tags:} [ECHO], [VEIL], [THRESHOLD]

\textbf{Threshold:} A deathbed or a grave.

Absorb the grief of one dying person. They die peacefully, forgetting their sorrow. You gain a Grief Token (counts as 1 Boon for any Rite involving memory or comfort). The grief is stored in your Cord.

\textbf{Obligation:} 1

\subsubsection{Low Rite: Rite of the Milk Teeth (4 XP)}

\textbf{Tags:} [BIND], [ECHO], [VEIL]

\textbf{Threshold:} A child's shed tooth, a drop of milk, a whispered promise.

Bind Reflection. The target's reflection is trapped in a mirror or pool. For one scene, the target acts at Position -1 (disoriented, seeing double) while the reflection reports their movements to you.

\textbf{Obligation:} 1

\subsubsection{Standard Rite: Rite of the Silent Tithe (7 XP)}

\textbf{Tags:} [PRICE], [THRESHOLD]

\textbf{Threshold:} A doorway of a house you have tended.

Once per season, you may declare that a household owes the Cord a tithe—a memory, a shadow, or a first breath of a newborn. The tithe is collected within the week, often without the household's awareness. You gain 1 permanent segment of Influence within that household's community.

\textbf{Obligation:} 2

\subsubsection{Standard Rite: Rite of the Flesh-Knot (8 XP)}

\textbf{Tags:} [BIND], [ECHO], [VEIL]

\textbf{Threshold:} A drop of your own blood, a whispered promise.

Store 1 Fatigue as a visible knot in your skin (a fresh scar, a weeping tattoo, a tightened cord). You may release the knot later to ignore 1 Fatigue from exertion. You may have up to Body rating knots at once.

\textbf{Obligation:} 2

\subsubsection{High Rite: Rite of the Hollowed Union (12 XP)}

\textbf{Tags:} [BIND], [NINTH]

\textbf{Threshold:} A birthing bed or a deathbed where three generations have passed.

Join two bloodlines through a debt that will be paid by a child not yet conceived. Write the terms in blood on a strip of linen, then feed the linen to your Hollowed. The Hollowed's eyes film over for a day. One year later, the debt will manifest—a stillbirth, a windfall, a sudden illness—as you decreed.

\textbf{Obligation:} 3

\subsection{Patron-Specific Talents}

\subsubsection{Cord of Flesh (8 XP)}

\textbf{Requirement:} The High Mother, Body 3+

Your body stores memories and debts. When you convert Harm → Fatigue (Monk ability), you may store up to Body rating points of Fatigue as Chained Knots, appearing uninjured (as fresh scars, weeping tattoos, or coiled threads beneath the skin). You have until the sunset after the next to perform a sacrifice or take 2 Harm.

\subsubsection{The Open Palm, Closed Cord (8 XP)}

\textbf{Requirement:} The High Mother, Melee 2+

Your unarmed strikes or subtle touch carry Widow's Measure. Instead of dealing Harm, you may mark a Promise Clock [4] on the target—they owe the Cord a truth or grief. If they already owe the coven, you may activate that debt: they suffer Position -1 for the scene, or you gain Effect +1 against them.

\subsubsection{The Red Thread (4 XP)}

\textbf{Requirement:} The High Mother

Once per scene, you may tie a small knot in a cord, thread, or cloth within reach. For the rest of the scene, you and allies within Near range gain +1 die to a single type of roll (choose when you tie the knot: stealth, persuasion, defense, or navigation). The knot must remain tied; if undone, the benefit ends.

\subsection{Roleplaying a Bloodsworn}

\begin{itemize}
\item You do not give gifts. You extend credit. The soup is free. The price never is.
\item You may be called to witness deathbed confessions, to tend the sick, or to collect debts that have compounded for generations.
\item You may struggle with the weight of what you carry. Every knot is a debt. Every scar is a payment. The Cord does not forget. Neither should you.
\item You may be asked to cut a thread—to release a debt, to forgive a promise, to let someone go free. The Cord does not forbid mercy. It asks only that you remember.
\end{itemize}

\subsection{Sample NPC: Sister Alva of the Unbroken Cord}

\textbf{Archetype:} Bloodsworn

\textbf{Patron:} The High Mother

\textbf{Codex:} Flesh-tattoos that weep when debts come due

\textbf{Thiasos:} A Hollowed child with milk-white eyes that never blink

\textbf{Personality:} Quiet, steady, inexhaustible. She has witnessed a thousand confessions. She has never repeated one.

\textbf{Conflict:} A debt she collected thirty years ago has come due again—not from the original debtor, but from their grandchild. The Cord does not distinguish between generations. Alva must decide: collect, or cut the thread for the first time.

\subsection{GM Guidance: Intrusions for The High Mother}

\begin{itemize}
\item A Hollowed's gaze lingers on a character. They feel watched. Next roll involving secrecy is Desperate.
\item A thread of grey cord appears tied around a character's wrist. They cannot recall when it happened. Removing it requires a test (DV 3) or they suffer -1 die to social rolls with anyone who knows the Cord.
\item A Daughter invokes a Threshold Contract the party unknowingly accepted (they ate soup at a Daughter's hospice). The party must choose: accept a minor debt or reject it (gain 2 SB banked for the Daughter in the next scene).
\item The High Mother's attention falls on a character. They dream of the Cord. When they wake, they have forgotten a significant memory (GM chooses—a face, a secret, a location).
\end{itemize}

The thread held. The water carried. That is not a Rite. That is grace.
