\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{parskip}
\usepackage{titlesec}

\definecolor{rust}{HTML}{8B3A3A}
\definecolor{stone}{HTML}{5A5A5A}
\definecolor{ink}{HTML}{1A1A1A}

\titleformat{\section}{\Large\bfseries\color{rust}}{}{0em}{}
\titleformat{\subsection}{\bfseries\color{stone}}{}{0em}{}

\begin{document}

\begin{center}
{\Large \textbf{A Player's Primer to the Amaranthine World}}
\end{center}

\vspace{-1em}

\begin{center}
\rule{0.8\textwidth}{0.5pt}
\vspace{0.5em}
\itshape\small
``The road does not speak in words. It speaks in suits, ranks, and the spaces between. 
Every bridge, every bribe, every broken oath can be laid out on a table of felt and dealt 
from a worn deck. This is not superstition. It is method.''
\vspace{0.3em}

--- Mira of the Wandering Circle, Tulkani cartomancer
\rule{0.8\textwidth}{0.5pt}
\end{center}

\vspace{1em}

\section*{How to Use This Handout}

This is not a rulebook. This is \emph{what your character knows}. Read it to understand:

\begin{itemize}
\item Where you come from and what that means
\item What you fear and why you fear it
\item What you carry and what it says about you
\item The one rule that everyone in the world obeys
\end{itemize}

You don't need to memorize this. You need to \emph{feel} it. When you sit down to play, you already know—without thinking—that the ninth cup is never poured, the ninth step is never taken, and the ninth bell is silence.

\section*{Before You Begin: The One Rule}

There is a law older than kings, written not in ink but in breath, in bells, in the space 
between heartbeats. It is the \textbf{Ninth Taboo}, and it touches every land in Fate's Edge.

\begin{quote}
\itshape
The wise stop at eight. They count eight breaths, ring eight bells, pour eight cups. 
The ninth is never taken. The ninth is the Hollow's share.
\end{quote}

You will see this everywhere:
\begin{itemize}
\item In \textbf{Vhasia}, the ninth citation in court voids all previous---and the dead king laughs.
\item In \textbf{Aeler}, counting to nine in the deep invites the pressure that thinks.
\item In \textbf{Aelaerem}, the ninth cup at a feast is for the empty chair that is never truly empty.
\item In \textbf{Theona}, the ninth step on any bridge is a missing step that leads to the Green Host.
\item In \textbf{Silkstrand}, the ninth bridge crossed in a day leads somewhere unexpected.
\end{itemize}

\textbf{What the Ninth Taboo Means for You:} Your character knows---\emph{knows}---that some 
numbers are not meant to be reached. You may not believe in gods, but you believe in the 
ninth step, the ninth word, the ninth toll. You have been taught this since childhood. 
You do not test it.

\section*{The Universal Truths}

Before you know where you're from, know what everyone knows. Across every land, these 
truths are as certain as sunrise:

\begin{itemize}
\item \textbf{The Hollow is patient.} It does not strike. It collects. A forgotten debt, a broken promise, a ninth cup poured—these are not sins. They are weights. The Hollow will come for them eventually. It waits until the debt is interesting.

\item \textbf{Names have weight.} To speak a true name is to give power. To forget a name is to leave a door unguarded. The dead remember their names. The living should be careful with theirs. A name spoken at a crossroads becomes a toll.

\item \textbf{Eight is the number of safety.} Eight breaths, eight bells, eight steps. The ninth is the Hollow's. The wise stop at eight. The wise know that the eighth breath is the last one before the Hollow counts.

\item \textbf{Hospitality is not kindness. It is a covenant.} A cup shared is a promise made. A threshold crossed is a bond formed. To break guest-right is to invite the Hollow to visit you next. A guest who eats three meals owes a service. A guest who eats nine owes a life.
\end{itemize}

\section*{Who Are You?}

You are a traveler, a survivor, a person with a past and a price. You come from one of 
the great regions of the Amaranthine world. Each has a flavor, a wound, and a way of 
looking at the world.

\subsection*{Kahfagia --- The Lantern Thalassocracy}

\textbf{Vibe:} Salt-spray politics, light as law, webs in the mirror-towers.

\textbf{You know:} Every beacon is a promise. Every mirror is a blade. The Admiralty 
controls the channels, but the \textbf{Lethai-ar} spiders weave through the oligarchs. 
You've seen silk threads on bollards and learned not to ask. The Mirror-Keepers are not 
born—they are made. The making requires a death.

\textbf{You carry:} A small brass bell—the sound of your name. Offering it means 
``I give you my voice.'' Refusing to return it is a declaration of blood feud.

\subsection*{Vhasia --- The Fractured Sun}

\textbf{Vibe:} Shattered chivalry, two crowns, the weight of Lence.

\textbf{You know:} The Fenwoods burned Lence ten years ago. The sun shattered, and 
the pieces are still falling. You've chosen a side, or you've refused to choose, but 
either way you remember the smoke. Chivalry is a mask. The knife is the truth. The 
Queen-Mother weaves alliances; the Two Crowns march; the Last Dauphin is a rumor with 
a birthmark.

\textbf{You carry:} A coin that bears two sunbursts—or one, depending on which 
mint you trust. The wrong metal on the wrong road is treason.

\subsection*{Thepyrgos --- The City of Unfinished Proofs}

\textbf{Vibe:} Vertical labyrinth, bell-schedules, the argument between memory and forgetting.

\textbf{You know:} The Synod rules with the delicacy of a brick wrapped in velvet. 
The \textbf{Chain-Lanterns} hunt untutored magic. The \textbf{Unfinished Stair} appears 
only to those who ask questions the Synod cannot answer. You've learned to count the 
bells, to watch your step, and to never ask the ninth question. The Lethai-thora 
refugees keep the deepest archives. Some serve the Witness; others whisper to the Ninth.

\textbf{You carry:} A stair-token—a bronze step that fits a missing rung. The 
missing rung is a secret.

\subsection*{Aeler --- The Crowns and Under-Vaults}

\textbf{Vibe:} Survival horror underground, breath-counting, the pressure that thinks.

\textbf{You know:} Air is measured. Water is bartered. Light is rationed. Every 
tunnel has a breaking point, and the deep remembers every breath you owe. The 
\textbf{Deep Drake} does not attack—it invites. You have learned to count your 
breaths, to ring your bell, and to never, ever answer a knock that comes in threes. 
The True Masons are not neutral—they are the Vault-Queen's eyes.

\textbf{You carry:} A breath-counter—a small brass bell. Eight breaths, ring the 
bell. Nine breaths, run.

\subsection*{Ykrul --- The Wolf Standards}

\textbf{Vibe:} Nomadic epic, Kon'reh geometry, the cold arithmetic of the steppe.

\textbf{You know:} The steppe owes nothing. The wolf pays no debts. You've learned 
Kon'reh—the game of encirclement and escape—before you could hunt. You count your 
exits before you enter. You never commit to a single path. The Khagan is chosen by 
the horse. The horse has not stepped forward in twenty years.

\textbf{You carry:} A hostage string—a braided cord that ties your wrist to 
another's. The cord is cut when the apology is accepted. If it is never cut, 
you are bound. The binding is called ``friend.''

\subsection*{Viterra --- The Hedge-Law Realm}

\textbf{Vibe:} Legal warfare, border disputes, the Queen's fragile peace.

\textbf{You know:} Every hedge is a boundary. Every boundary is a claim. Every claim 
is a blade. The Queen's peace is a contract written in ink and witnessed by steel. 
You've learned to read a charter, to know which seal is warm, and to never cross 
the Belworth without papers recognized on both banks. The Queen's crown has no jewels—only seals from conquered claims.

\textbf{You carry:} A seal ring—worn thin from constant gripping. Your signature 
is your bond. Your bond is your life.

\subsection*{Mistlands --- Bells, Salt, and Breath}

\textbf{Vibe:} Folk horror, fog-bells, salt as law, the Direwood's breath.

\textbf{You know:} The bell-lines keep the wraiths back. The ward-salt is the only 
thing that stops the dead from rising. The Aeler Protectorate maintains order, but 
the old ways are older. You know to ring eight bells, to never answer a voice that 
calls from the fog, and that the ninth bell summons the Hollow. The Legate of the 
Mists is a wraith, rebuilt from fog and memory. She does not know she is dead.

\textbf{You carry:} A salt-pouch—measured and blessed. Spill it, and the fog remembers.

\subsection*{Linn --- Skerries and Storm-Oaths}

\textbf{Vibe:} Norse coastal saga, frost-bound oaths, the tyranny of seasons.

\textbf{You know:} The sea keeps better counsel than any scribe. The frost is the 
witness—it preserves what the tide would wash away. A promise sworn in winter is 
owed until the ice melts. You've learned to tie the eight sea-knots. The ninth 
knot is for emergencies—and for funerals.

\textbf{You carry:} An oath-stick—a carved stick, each notch a promise kept or broken. 
The frost remembers.

\subsection*{Ubral --- The Stone Between Spears}

\textbf{Vibe:} Highland clan honor, wergild, the geometry of grudges.

\textbf{You know:} The stone remembers every oath sworn upon it. The dead vote 
at the cairns. A guest's word can buy more than a sword, but a broken vow sparks 
feuds lasting generations. You've learned to count in cattle and blood-price. 
The ninth stone in any cairn is always warm.

\textbf{You carry:} A guest-token—a knucklebone from a feast. The host's name 
is carved on it. Break it to end the obligation.

\subsection*{Silkstrand --- City of Bridges and Dyewater}

\textbf{Vibe:} Mercantile intrigue, layered loyalties, the Ragman's threads.

\textbf{You know:} Every bridge is a promise. Every canal carries a secret. The 
old patrician families—Contarini, Morosini, Dandolo—hold charters signed by dead 
emperors. The Ragman holds the alleys. The Matron balances the threads. A favor 
is a thread that can be pulled from either end.

\textbf{You carry:} A favor thread—red for a promise owed, black for a promise 
owed to you. The ninth thread is the Ragman's. He will find you.

\subsection*{Valewood --- Empire Under Leaves}

\textbf{Vibe:} Fairy tale, dark forest, fey remnant, the soft place.

\textbf{You know:} The forest remembers empires that crumbled to dust. The 
Genesis Vats still hum beneath the Amber Ziggurat. The fae courts of Hazel 
and Alder play their endless games. The beast-kin are the descendants of 
survivors. You've learned to speak truth, to step lightly, and to never eat 
the fruit of a tree with a face.

\textbf{You carry:} A name-bead—a bead that warms when you keep a promise 
and chills when you break one. If it glows, the forest has heard you.

\subsection*{Black Banners --- Condotta and Crowns}

\textbf{Vibe:} Mercenary realism, logistics-driven violence, loyalty in the mud.

\textbf{You know:} A contract is only as good as the next payday. A grudge is 
only as sharp as the sword backing it. You've learned to count coin, to know 
which colors fly for which captain, and to never break a Bell Dawn. The Black 
Colonel is not a person—the title is a wax seal in a lead-lined box.

\textbf{You carry:} A pay-tally—a notched stick. Each notch is a day served. 
The ninth notch means you've earned the right to demand a story.

\section*{If Your Region Isn't Listed}

You are not limited to these regions. The Amaranthine world has many lands:

\begin{itemize}
\item \textbf{Acasia} — Broken Marches, the Curse, free companies, the Pale Causeway
\item \textbf{Aelaerem} — Halfling downs, hearth magic, the Pale Shepherd, the Hollow
\item \textbf{Aelinnel} — Gnomish courts, fae bargains, the Green Gate, oath-beads
\item \textbf{Theona} — Three Greens, the No Ninth taboo, the Green Host, the Lady Beneath
\item \textbf{Dungeons} — Living infrastructure, forgotten machinery, the pressure below
\item \textbf{The Wilds} — Hexcrawl wanderlust, weather as character, the land's patience
\item \textbf{The Ways Between} — Dream-death metaphysical, virtue and vice, the Thirteenth Stone
\end{itemize}

Ask your GM. Each region has its own rules, its own fears, and its own tokens. 
You are not limited to these pages. You simply need to know what you carry 
and what you owe.

\section*{The World at a Glance}

\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Region} & \textbf{Tagline} & \textbf{What You'll Face} \\
\midrule
Kahfagia & ``Every mirror is a blade, every beacon a promise'' & Lantern-politics, infiltration, reefs \\
Vhasia & ``The sun shattered at Lence. The pieces are still falling'' & Court intrigue, fractured loyalty, chevauchée raids \\
Thepyrgos & ``Every stair is a question. Every answer is a proof'' & Bell-schedules, Synod law, hidden stairs \\
Aeler & ``The mountain counts your breaths'' & Survival, breath-counting, the pressure below \\
Ykrul & ``The steppe owes nothing. The wolf pays no debts'' & Nomadic honor, Kon'reh, hostage strings \\
Viterra & ``Every hedge is a boundary. Every boundary is a claim'' & Legal warfare, border disputes, charters \\
Mistlands & ``The fog does not forget. It remembers every name'' & Bell-lines, wraiths, salt as law \\
Linn & ``The frost is the witness. The sea is the judge'' & Storm-oaths, winter debts, longship raids \\
Ubral & ``The stone remembers. The dead vote'' & Clan feuds, wergild, cairn law \\
Silkstrand & ``Every bridge is a promise. Every canal carries a secret'' & Mercantile intrigue, layered loyalties, threads \\
Valewood & ``The forest remembers empires that crumbled to dust'' & Fae courts, beast-kin, sealed vats \\
Black Banners & ``A contract is only as good as the next payday'' & Mercenary warfare, logistics, camp politics \\
\bottomrule
\end{tabular}

\section*{The Three Voices}

This world is told through three travelers who disagree. They all walked these roads. 
They all paid their tolls. Listen to each---they will contradict each other, and that 
is the point.

\subsection*{Lady Isolde of the Scar}
\emph{``Chivalry is not a code---it is a weight. You carry it or it crushes you.''}

She speaks for Vhasia, for Viterra, for the hedge-law courts and the fractured sun. 
She counts stones. She sees the crack in every seal. Trust her when she warns you 
of a lord. She has seen the knife beneath the cloak.

\subsection*{The Bravo of Silkstrand}
\emph{``The best secrets are the ones you share over wine.''}

He speaks for Silkstrand, for the Brass Coast, for the canals that carry secrets. 
He has won purses and lost friends and still smiles. Trust him when he tells you 
a bridge will hold. He has counted every plank.

\subsection*{The Edgewalker}
\emph{``The mountain breathes. The stone remembers. I am the crack in both.''}

He speaks for Aeler, for the Mistlands, for the deep vaults where the Hollow is 
not a myth but a flaw in the foundation. Trust him when he tells you to flee. 
The stone's memory is longer than yours.

\section*{For the Rules}

This handout is the world you live in. For the rules of play—how to roll dice, what 
Boons are, how Position works—see the \textbf{Player Cyclopedia}. It's one page. 
You can read it in under two minutes.

\section*{Getting Started}

Before your first session, know these three things about your character:

\begin{enumerate}
\item \textbf{What do you owe?} A debt, a promise, a broken oath---something you 
haven't paid yet. It doesn't have to be large. It just has to be real.

\item \textbf{What do you carry?} A bell, a seal, a coin, a thread, a name. 
Something that reminds you of where you came from and what you've left behind.

\item \textbf{What do you fear?} The ninth bell. The empty chair. A knock in threes. 
Something that every child in your homeland was taught to fear---and you still do.
\end{enumerate}

\vspace{1em}
\begin{center}
\rule{0.6\textwidth}{0.5pt}
\vspace{0.5em}
\itshape\small
``The road is never closed. It is only passed to the next walker.''
\vspace{0.3em}

--- Mira of the Wandering Circle, on the eve of a journey she did not survive
\rule{0.6\textwidth}{0.5pt}
\end{center}

\newpage

\section*{Read Me First}

Copyright \copyright\ Nicholas A. Gasper

First Edition, June 2026

\vspace{0.5cm}

\textbf{All Rights Reserved.}

This work is the intellectual property of Nicholas A. Gasper and may not be reproduced, distributed, or transmitted in any form or by any means without prior written permission, except for brief quotations in reviews and certain other noncommercial uses permitted by copyright law.

\vspace{0.3cm}

The Fate's Edge System Reference Document (SRD), as included in this work, is licensed separately under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License. The SRD content is clearly marked throughout this book. All other content is proprietary and subject to the All Rights Reserved notice above.

\vspace{0.3cm}

For permission requests or licensing inquiries, contact the author through the publisher's website or at the address provided in the commercial edition.

\vspace{1cm}

\begin{center}
\rule{0.4\textwidth}{0.5pt}

\textit{``The coin that never spends is the one you don't remember taking.''}

--- Serafine of the Velvet Touch

\rule{0.4\textwidth}{0.5pt}
\end{center}

\vspace{0.5cm}

\subsection*{SRD Designation}

This document contains material designated as the Fate's Edge System Reference Document (SRD). The SRD is the open-license portion of this work and is intended for use by third-party creators, developers, and community members.

\subsubsection*{SRD Content Includes:}
\begin{itemize}
\item Core resolution mechanics (dice pools, Position, Difficulty Value, Outcome Matrix)
\item Boon and Story Beat economies
\item Fatigue, Harm, and Armor Conversion systems
\item Attribute and Skill descriptions
\item Basic combat and action economy rules
\item Universal magic mechanics (TAGS, DV calculation, Backlash)
\item NPC templates and encounter guidelines
\item Equipment and gear tables
\item Character creation framework
\item Optional systems (Deck of Consequences, Crown Spread, Travel framework)
\end{itemize}

\subsubsection*{Extended Content (Not SRD):}
\begin{itemize}
\item Setting lore, regional descriptions, and cultural details
\item Original characters, NPCs, and named figures
\item Adventure modules and scenario hooks
\item Patron descriptions, Rites, and exclusive mechanical content
\item Proprietary magic systems (Runkeeper, Invoker, Cantor, Summoner-specific expansions)
\item Artwork, maps, and graphical elements
\item Faction descriptions and campaign-specific content
\item Original prose, framing devices, and narrative text
\end{itemize}

\subsubsection*{License Summary:}

The SRD content in this document is licensed under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License.

All extended content is All Rights Reserved and the intellectual property of Nicholas A. Gasper.

For the full license text, see the copyright page.

\vspace{0.5cm}

\begin{center}
\rule{0.4\textwidth}{0.5pt}

\textit{``The coin that never spends is the one you don't remember taking.''}

--- Serafine of the Velvet Touch

\rule{0.4\textwidth}{0.5pt}
\end{center}

\input{copyright}
\end{document}