% ============================================================
% Fate's Edge --- Complete Worldbook
% Revised with layered narration and regional generators
% ============================================================
\documentclass[11pt,letterpaper,twoside]{book}

% ---------- Encoding, fonts, micro-typography ----------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[final]{microtype}

% ---------- Page geometry ----------
\usepackage[letterpaper,margin=1in]{geometry}

% ---------- Math & symbols ----------
\usepackage{amsmath,amssymb,mathtools}

% ---------- Colors, graphics, layout ----------
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning}
\usetikzlibrary{calc,shadows.blur,decorations.pathmorphing}

% ---------- Tables ----------
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{longtable}

% ---------- Lists ----------
\usepackage{enumitem}
\setlist{topsep=4pt,itemsep=2pt,parsep=0pt,partopsep=0pt}
\setlist[description]{style=nextline,leftmargin=2.2em,labelsep=0.6em}

% ---------- Hyperlinks & references ----------
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=[rgb]{0.10,0.25,0.55},
  citecolor=[rgb]{0.10,0.45,0.25},
  urlcolor=[rgb]{0.55,0.10,0.10},
  pdfauthor={Fate's Edge},
  pdftitle={Fate's Edge --- Complete Worldbook},
  pdfcreator={LaTeX}
}
\usepackage{bookmark}
\usepackage[nameinlink,noabbrev,capitalise]{cleveref}

% ---------- Indexing ----------
\usepackage{imakeidx}
\makeindex

\newcommand{\indexterm}[2][]{%
  \textit{#2}%
  \if\relax\detokenize{#1}\relax\index{#2}\else\index{#1}\fi%
}

% ---------- Table helpers ----------
\renewcommand{\arraystretch}{1.15}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\definecolor{FeRow}{HTML}{F5F7FA}
\newcommand{\feTableStart}{\rowcolors{2}{FeRow}{white}\small}
\newcommand{\feTableEnd}{\rowcolors{2}{}{}}

% ---------- Small conveniences ----------
\newcommand{\textrarr}{\ensuremath{\rightarrow}}
\newcommand{\textlarr}{\ensuremath{\leftarrow}}
\newcommand{\textdarr}{\ensuremath{\downarrow}}
\newcommand{\beat}{\ensuremath{\triangleright}\,}
\newcommand{\fullsquare}{\raisebox{0.2ex}{\rule{0.9ex}{0.9ex}}}
\newcommand{\Tag}[1]{\texttt{[#1]}}

% ---------- Suit symbols ----------
\newcommand{\SuitSpade}{\ensuremath{\spadesuit}}
\newcommand{\SuitHeart}{\ensuremath{\heartsuit}}
\newcommand{\SuitClub}{\ensuremath{\clubsuit}}
\newcommand{\SuitDiamond}{\ensuremath{\diamondsuit}}

% ---------- Almanac environment ----------
\newtcolorbox{almanac}{
  colback=gray!5,
  colframe=fePrimary,
  fonttitle=\bfseries\large,
  title={\textsc{The Almanac of Ordinary Days}},
  sharp corners,
  boxrule=1pt,
  left=12pt,
  right=12pt,
  top=8pt,
  bottom=8pt
}

% ---------- Branding ----------
\definecolor{fePrimary}{HTML}{0D1B2A}
\definecolor{feAccent}{HTML}{6FFFE9}
\definecolor{feInk}{HTML}{101418}
\definecolor{fePaper}{HTML}{F7F7F7}

\newcommand{\feTitlePage}{%
  \begin{titlepage}
    \centering
    \vspace*{3cm}
    {\Huge\bfseries Fate's Edge \par}
    \vspace{1cm}
    {\LARGE Complete Worldbook \par}
    \vspace{0.5cm}
    {\large \itshape A Gazetteer of the Known Realms\par}
    \vfill
    {\large \textsc{As told by those who live there, bleed there, and owe there}\par}
    \vspace{1cm}
  \end{titlepage}
}

\begin{document}

\feTitlePage
\tableofcontents

% ============================================================
% PREFACE: THE LEDGER BLEEDS
% ============================================================
\chapter*{Preface: The Ledger Bleeds}
\addcontentsline{toc}{chapter}{Preface: The Ledger Bleeds}
\label{chap:preface}

\noindent\textit{I am an Aeler Edgewalker. You call us mountain-dwellers who measure breath in silver bells and seal contracts in stone. How adorably naïve.}

We don't just measure breath—we own the lungs that draw it. When a Tulkani chieftain swears an oath on his firstborn's life? That's not honour—it's collateral I've already hypothecated three times over. When a Veiled Sister binds a demon to a corpse to prop up her corpse-throne? That's not terror—it's a futures contract on soul-debt, and I'm the broker who sold the short position before the ritual even began.

You see the surface as "full of problems." How quaint. Problems aren't obstacles—they're inventory.

This book isn't a tourist's guide. It's a hit list with footnotes. I've marked the places where a well-placed bell doesn't just open a pass—it starts a war I've already shorted the bond market on. Where a forged seal doesn't just move a mountain—it triggers a landslide I've bought mineral rights to before the first crack appeared. Where a single true story doesn't just buy a ship's loyalty—it buys the captain's silence while I sell his manifest to pirates.

You want to build journeys? Draw your cards. Set your timers. Spend your Story Beats like they mean something. But remember: every Diamond you earn is a tooth knocked out of someone else's jaw. Every Debt you incur? I already bought it from your victim for half what you'll pay—and I'm collecting interest in your marrow before you finish reading this sentence.

The road isn't just a ledger. It's a butcher's block. Every gate isn't a clause—it's a throat waiting for the knife.

\vspace{1em}
\noindent\textit{— The Edgewalker}

\newpage

\chapter*{The Cartomancer’s Compass}
\label{chap:cartomancer-compass}

\section*{The Ledger of the Deck}

The world does not speak in words. It speaks in suits, ranks, and the spaces between.
Every bridge, every bribe, every broken oath can be laid out on a table of felt and
dealt from a worn deck. This is not superstition. It is method.

The regional generators in this book \— from the fog-choked levees of the
Mistlands (\autoref{chap:mistlands}) to the walking cities of Ngomebe
(\autoref{chap:ngomebe}) \— all speak the same language: the language of the
fifty-two-card deck. Learn to read it, and you can generate a campaign, a
crossroads, or a crisis in the time it takes to shuffle.

This chapter is your key. It explains how to interpret any card drawn from any
regional generator, and how to weave those interpretations into the fiction at
your table. Consider it the threshold between the abstract mechanics of the draw
and the living, breathing world of Fate’s Edge.

\section*{The Four Suits: A Universal Grammar}

Every card in every generator carries the same four suit-meanings. They are the
alphabet of the world’s ledger.

\subsection{\texorpdfstring{\spadesuit}{Spades}: Place \— The Stage}
\label{sec:spades-place}

Where does the scene happen? \spadesuit\ is the threshold, the landmark, the ground
beneath the action. It answers: \emph{Where are we?}

\begin{itemize}
    \item \textbf{High rank (10–A)}: A major landmark \— a fortress, a capital, a
    haunted ruin. The place has a name that echoes in local songs.
    \item \textbf{Mid rank (6–9)}: A functional location \— a ford, a market, a
    shrine. The place has a purpose and a keeper.
    \item \textbf{Low rank (2–5)}: A liminal or neglected place \— a crossroads, a
    collapsed bridge, a smuggler’s cove. The place is defined by what it lacks.
\end{itemize}

In the regional chapters, \spadesuit\ entries are always places. See, for example,
the bridges and canals of Silkstrand (\autoref{sec:silkstrand-places}) or the
cairns and passes of Ubral (\autoref{sec:ubral-places}).

\subsection{\texorpdfstring{\heartsuit}{Hearts}: Actor \— The Face}
\label{sec:hearts-actor}

Who drives the scene? \heartsuit\ is the person, the faction, the voice that
speaks. It answers: \emph{Who is here?}

\begin{itemize}
    \item \textbf{High rank (J–A)}: An elite, a ruler, a power-broker. The actor has
    authority and reach beyond the immediate scene.
    \item \textbf{Mid rank (6–10)}: A professional, a functionary, a skilled
    operator. The actor has a trade and a price.
    \item \textbf{Low rank (2–5)}: A commoner, an outcast, a victim. The actor
    needs something or has been failed by someone.
\end{itemize}

In the regional chapters, \heartsuit\ entries are always people or factions. See the
merchants and spies of the Brass Coast (\autoref{sec:brasscoast-people}) or the
spirit-wardens of the Mistlands (\autoref{sec:mistlands-people}).

\subsection{\texorpdfstring{\clubsuit}{Clubs}: Pressure \— The Complication}
\label{sec:clubs-pressure}

What pushes back? \clubsuit\ is the storm, the debt, the knife in the dark. It
answers: \emph{What goes wrong?}

\begin{itemize}
    \item \textbf{High rank (10–A)}: A catastrophe, a betrayal, an invasion. The
    complication changes the shape of the journey.
    \item \textbf{Mid rank (6–9)}: A serious obstacle \— a feud, a quarantine, a
    broken bridge. The complication costs time, resources, or reputation.
    \item \textbf{Low rank (2–5)}: A minor hazard \— a sudden squall, a lost
    token, a raised toll. The complication is a nuisance, not a disaster.
\end{itemize}

In the regional chapters, \clubsuit\ entries are always complications or threats.
See the river floods and witness-cord cuts of Ayokha
(\autoref{sec:ayokha-complications}) or the market riots of Silkstrand
(\autoref{sec:silkstrand-complications}).

\subsection{\texorpdfstring{\diamondsuit}{Diamonds}: Leverage \— The Reward}
\label{sec:diamonds-leverage}

What opens the door? \diamondsuit\ is the token, the writ, the secret that changes
Position. It answers: \emph{What do you gain?}

\begin{itemize}
    \item \textbf{High rank (J–A)}: A major asset \— a noble title, a keystone
    charter, a general pardon. The reward reshapes the party’s standing.
    \item \textbf{Mid rank (6–10)}: A significant tool \— a pass-stone, a witness
    cord, a lighthouse charter. The reward solves a specific class of problem.
    \item \textbf{Low rank (2–5)}: A minor advantage \— a bridge token, a guest-cup,
    a rumour. The reward grants a single use of improved Position.
\end{itemize}

In the regional chapters, \diamondsuit\ entries are always rewards or leverage. See
the keystone charters of Ngomebe (\autoref{sec:ngomebe-rewards}) or the
water-share tablets of Fhara (\autoref{sec:fhara-rewards}).

\section*{Rank and Severity: The Weight of the Card}
\label{sec:rank-severity}

The rank of a card \— from Ace (high) to 2 (low) \— sets the scale of its suit’s
meaning. Use this ladder:

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Rank} & \textbf{Narrative Weight} \\
\hline
Ace & Legendary, campaign-altering, mythic \\
King/Queen/Jack & Major, faction-level, lasting consequences \\
10–7 & Significant, scene-defining, costly \\
6–4 & Moderate, manageable, a real cost \\
3–2 & Minor, a complication but not a crisis \\
\hline
\end{tabular}
\end{center}

When a generator instructs you to set a timer by the highest rank, use this
conversion:

\begin{itemize}
    \item \textbf{2–5}: 4-segment timer (short, local problem)
    \item \textbf{6–10}: 6-segment timer (standard, journey-length)
    \item \textbf{J, Q, K}: 8-segment timer (extended, regional problem)
    \item \textbf{A}: 10-segment timer (epic, campaign-length)
\end{itemize}

\section*{Card Faces: The Court as Character}
\label{sec:court-characters}

Face cards (Jack, Queen, King) are not merely high ranks. They are \emph{persons}
with agency, personality, and often a name.

\begin{itemize}
    \item \textbf{Jack}: An agent, a functionary, a messenger. The Jack acts on
    someone else’s authority. They are competent but replaceable.
    \item \textbf{Queen}: A power in their own right. The Queen has influence,
    resources, and a network. They are not easily replaced.
    \item \textbf{King}: The ultimate authority in their domain. The King is the
    source of law, the peak of the hierarchy. To challenge a King is to challenge
    the order itself.
\end{itemize}

When a face card appears in a \heartsuit\ (actor) draw, give the NPC a name, a
motive, and a quirk. When it appears in a \spadesuit\ (place) or \clubsuit\
(pressure) draw, treat it as a \emph{personified} force: a King of Clubs might be
a feud that has lasted generations; a Queen of Spades might be a bridge that
remembers every toll.

\section*{Aces and the Ninth: The Hollow’s Attention}
\label{sec:aces-ninth}

An Ace is not merely the highest rank. It is a \emph{threshold}. When an Ace
appears in any draw, the Hollow takes notice. The GM gains 1 extra Story Beat
(beyond any generated by 1s) to spend on a supernatural complication: a ghostly
procession, a bell that rings without a clapper, a shadow that walks
independently.

In regions where the Ninth Taboo is strong (see \autoref{sec:theona-mechanics},
\autoref{sec:aelinnel-mechanics}, \autoref{sec:ways-between-features}), an Ace may
also trigger a specific regional omen. For example:

\begin{itemize}
    \item In Theona (\autoref{chap:theona}), an Ace of \spadesuit\ might add a
    “missing step” to a bridge.
    \item In the Ways Between (\autoref{chap:ways-between}), an Ace of \heartsuit\
    might cause a reflection to speak.
    \item In the Mistlands (\autoref{chap:mistlands}), an Ace of \clubsuit\ might
    cause a bell-line to fail.
\end{itemize}

The GM should always read an Ace as a signal to escalate the uncanny.

\section*{By the Numbers: Procedure for Any Generator}
\label{sec:generator-procedure}

When you sit down to generate a scene, a journey leg, or a settlement, follow this
procedure:

\begin{enumerate}
    \item \textbf{Name the destination}. Which region? Which route? Which authority
    controls the gate?
    \item \textbf{Shuffle and draw}. Draw cards face-up, one for each suit you
    need. Most generators require all four suits; some (like travel legs) may use
    only three.
    \item \textbf{Read the \spadesuit}. This is your place. Describe it first. Set
    the stage.
    \item \textbf{Read the \heartsuit}. This is your actor. Introduce them. What do
    they want? Why are they here?
    \item \textbf{Read the \clubsuit}. This is your pressure. Complicate the scene.
    What just went wrong, or what is about to?
    \item \textbf{Read the \diamondsuit}. This is your leverage. What can the
    players gain if they succeed? Be specific.
    \item \textbf{Set the timer}. Use the highest rank among the four cards to
    determine the size of the scene’s primary timer (see \autoref{sec:rank-severity}).
    \item \textbf{Interpret faces and Aces}. Give names to face cards. Apply
    regional Ace effects.
    \item \textbf{Play to find out}. Use the generated elements as prompts, not
    scripts. The cards are seeds, not fences.
\end{enumerate}

\section*{Regional Flavours: How Suits Shift Meaning}
\label{sec:regional-flavours}

While the core suit meanings remain constant, each region inflects them with its
own texture. A \spadesuit\ in Ecktoria (\autoref{chap:ecktoria}) is a forum or a
coin-house; a \spadesuit\ in the Valewood (\autoref{chap:valewood}) is a phasing
ruin or a star-road. A \heartsuit\ in the Black Banners
(\autoref{chap:black-banners}) is a mercenary captain or a condotta broker; a
\heartsuit\ in Ykrul (\autoref{chap:ykrul}) is a khatun or a bone-singer.

Use the regional chapters as your lexicon. When you draw a card, consult the
relevant table (e.g., \autoref{sec:acasia-places} for Acasia, \autoref{sec:oshiira-people}
for Oshiira). The table will give you a specific entry keyed to the rank and suit.
Do not invent from scratch unless the table runs out \— the tables are designed to
evoke the specific flavour of the region.

\subsection*{Example: A Draw in Acasia}
\label{sec:example-acasia}

You are running a journey through the Broken Marches (\autoref{chap:acasia}). You
draw:

\begin{center}
\spadesuit\ 7, \quad \heartsuit\ \text{J}, \quad \clubsuit\ 4, \quad \diamondsuit\ \text{Q}
\end{center}

\begin{enumerate}
    \item \textbf{\spadesuit\ 7}: You consult \autoref{sec:acasia-places}. A 7 of
    Spades is “Hill-Motte \— Fresh palisade; ditch scattered with caltrops.”
    \item \textbf{\heartsuit\ J}: You consult \autoref{sec:acasia-people}. A Jack of
    Hearts is “Bride with No Dowry but a Claim — Daggers in the wedding chest.”
    \item \textbf{\clubsuit\ 4}: You consult \autoref{sec:acasia-complications}. A
    4 of Clubs is “Wedding Turns Ambush — You are caught between two bloodlines.”
    \item \textbf{\diamondsuit\ Q}: You consult \autoref{sec:acasia-rewards}. A Queen
    of Diamonds is “Marriage Proxy — From the Margravine; bind a hillking to your cause.”
    \item \textbf{Highest rank}: Queen (rank equivalent to King/J/Q at 8). The scene’s
    primary timer is 8 segments: \textit{The Feud Escalates [8]}.
    \item \textbf{Interpretation}: The party arrives at a hill-fort (Hill-Motte) on
    the eve of a wedding. The bride (Bride with No Dowry) is caught between two
    feuding bloodlines. The wedding turns to violence (Wedding Turns Ambush). If
    the party can survive the ambush and broker peace, the Margravine will grant
    them a Marriage Proxy — a powerful tool to bind a hill-king to their cause.
    The 8-segment timer tracks how quickly the feud spirals into open war.
\end{enumerate}

This is the generator at work: four cards, one scene, immediate fiction.

\section*{The Unwritten Suit: When a Draw is Incomplete}
\label{sec:incomplete-draw}

Sometimes, a generator does not require all four suits. A travel leg, for example,
may only need \spadesuit, \heartsuit, \clubsuit\ (see
\autoref{sec:globaltravel-procedure}). A social encounter might only need
\heartsuit\ and \clubsuit. When a suit is missing, treat its absence as a
question:

\begin{itemize}
    \item \textbf{No \spadesuit}: The place is ambiguous, shifting, or hidden. The
    scene occurs in a liminal space — a crossroads, a fog-bank, a threshold.
    \item \textbf{No \heartsuit}: The actor is absent, silent, or dead. The
    pressure and place act without a face. This can be more frightening than a
    named NPC.
    \item \textbf{No \clubsuit}: The scene is calm, peaceful, or mercifully
    straightforward. This is rare in Fate’s Edge. Treasure it.
    \item \textbf{No \diamondsuit}: The reward is not a token but an outcome. The
    players gain Position, information, or a cleared timer rather than a physical
    asset.
\end{itemize}

\section*{Interpreting Any Draw, Any Region}}
\label{sec:any-draw}

When you need a prompt in a hurry — a random encounter, a rumour at a tavern, a
discovery in a ruin — draw a single card. Use this quick-reference:

\begin{center}
\begin{tabular}{|l|l|}
\hline
\spadesuit & A place or landmark. What is its name? What happened here? \\
\heartsuit & A person or faction. What do they want? What do they fear? \\
\clubsuit & A complication or threat. Who or what is the obstacle? \\
\diamondsuit & A reward, token, or secret. What can be gained? \\
\hline
\end{tabular}
\end{center}

The rank tells you the scale. A low rank is local and personal; a high rank is
regional and political. A face card demands a name; an Ace demands the Hollow’s
attention.

This single-card method is ideal for improvisation. It gives you one clear element
to build around. Let the fiction answer the rest.

\section*{Final Guidance: The Cards Are Oracles, Not Scripts}
\label{sec:cards-oracles}

The generators in this book are tools, not tyrants. If a draw gives you an entry
that contradicts established fiction, or if you simply prefer a different
interpretation, overrule it. The deck is a prompt, not a prison.

But trust it, too. The most memorable sessions often come from the card you did
not expect. The Ace that arrives at the wrong moment. The Jack who speaks a name
no one wanted to hear. The diamond that glows with a secret debt.

Shuffle. Draw. Interpret. And let the road unfold.

\vspace{1em}
\textit{“The ledger is never closed. It is only passed to the next keeper.”}
\hfill — Dusana of the Raven Road

\newpage
% ============================================================
% SNIPPET 3: GM Guidance Sidebar – Running the Omen Lore
% Suggested placement: In the front matter (after the Introduction) or as a sidebar in the GM's Appendix above.
% ============================================================

\begin{tcolorbox}[colback=black!2,colframe=blue!60!black,title={For the GM – The Spider, the Snake, the Raven}]
  The three Omens (\textbf{Inaea}, \textbf{Isoka}, \textbf{Ikasha}) are not gods. They are forces that appear in whispers, dreams, and the spaces between counted things. Do not introduce them directly. Instead, let players encounter their echoes:
  \begin{itemize}
    \item A rope that never frays (Inaea)
    \item A judgment that cuts clean (Isoka)
    \item A dream that feels too real (Ikasha)
  \end{itemize}
  By the time players learn the names, they should have already felt their presence for many sessions.

  The full truth of the Omens – the Timer work Demon, the scouring of civilisations, Ikasha’s imprisonment in Obishaal – is \textbf{endgame content}. Use the fragments in this book as breadcrumbs:
  \begin{itemize}
    \item If players ask a Pereshi fire‑keeper about the “serpent judge,” they will receive a riddle.
    \item If they ask a Tulkani about the “dreamer,” they will receive a story.
    \item If they ask a Sidhi about the “spider,” they will receive a knot and a warning: \textit{“Do not pull it unless you are ready to see what it holds.”}
  \end{itemize}
  The Hollow is the interest on the Omens’ oldest debt. The players are never meant to pay it off – only to decide whose name goes on the next line of the ledger.
\end{tcolorbox}
\newpage
% ============================================================
% PART I: THE AMARANTHINE SEA & NORTHERN SHORES
% ============================================================
\part{The Amaranthine Sea \& Northern Shores}

% After \part{The Amaranthine Sea & Northern Shores}
\begin{quote}
  \small
  \textit{``The sea keeps a better ledger than any scribe. The reef remembers every secret whispered above its depths. Do not ask which flag flies highest – ask which hull leaks slowest.''}
  \noindent — Dusana of the Raven Road
  \end{quote}

% ------------------------------
% CHAPTER: KAHFAGIA
% ------------------------------
\chapter{Kahfagia}
\emph{The Flag of Storms.}

\noindent
{\Large
\textit{I was born on the tide, which is to say I was born owing. My mother was a net-mender who lost her fingers to a reef-cutter's debt. My father was a bell—that is to say, he rang when the Admiralty pulled his string. I learned to read the lantern codes before I learned my letters, and I learned that every light is a lie waiting to be told.}

\vspace{0.5em}

The Admiralty runs this place like a pirate charter with a budget—and oh, how they love pretending it's legitimate. They control the beacons that lure ships onto reefs, the booms that shatter hulls for salvage rights, and the tariffs that fund their champagne baths. But here's the secret the commodores don't want you to know: the lanterns are afraid of the dark too.

\vspace{0.5em}

I've seen a Mirror-Keeper weep when her reflection showed her a daughter she'd sold for a pilot's license. I've watched a Privateer Captain kiss a silver coin before a raid—the same coin her mother kissed before she never came back. This is not a city of heroes. It is a city of people who learned that the sea is cheaper than mercy.

\vspace{0.5em}

If you want to move cargo through the straits without becoming fish food, get a convoy letter stamped with the Admiral's seal. Or get smart. Or get gone.

\vspace{1em}

\noindent\textit{— Captain Mira ``the Lucky,'' Sidhi exile, pirate by convenience}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salt fish porridge—dried cod soaked overnight, boiled with hardtack and a splash of lamp oil. The oil is for calories. The fish is for memory. Every Kahfagian child learns to taste the salt and know which coast their grandmother came from.

\textbf{What does a child learn before they can walk?} To count ships by their lanterns. Three lanterns means a Kahfagian merchant—safe. Four means an Oshiiran grain barge—neutral. One, low in the water, means a privateer. Run.

\textbf{What is the first thing you notice about a stranger?} Their hands. Dock-workers have rope-burned palms. Merchants have ink-stained fingers. Privateers have no fingernails—they bite them off in storms.

\textbf{What is the most common crime?} False manifests. Every third ship carries cargo that doesn't appear on its papers. The Admiralty knows. The Admiralty takes a cut. The Admiralty calls it a "tariff surcharge."

\textbf{What does no one talk about but everyone knows?} The Mirror-Keepers are not born—they are made. And the making requires a death. A sister, a daughter, a lover. Someone who drowned so the keeper could see through the fog.

\textbf{Where do people go when they die?} The tide takes them. Kahfagia has no cemeteries above ground. The drowned are weighted with chain and dropped in the deep channel. Their families light candles in paper boats and float them out at dusk. If the candle reaches the horizon, the soul is forgiven. If it sinks, the debt is not yet paid.

\textbf{How do you apologize for a serious wrong?} You offer your bell. Every Kahfagian carries a small brass bell on a leather cord—the sound of their name. Offering it to the wronged party means "I give you my voice. Speak for me." Refusing to return the bell after an apology is a declaration of blood feud.

\textbf{What is the most important festival?} The Beacon Lighting (midwinter). Every lighthouse keeper lights a new lamp from the Admiral's flame. If your beacon catches first, your year will be lucky. If it sputters, you will owe a death to the sea.

\textbf{What is the secret that could destroy a powerful person?} The Admiral's lantern code has a ninth signal—never used, never spoken, burned into every keeper's memory. It means \textit{"Kill the witnesses. Sink the evidence. The sea will hide the rest."}
\end{almanac}

\newpage
\input{../generators/kahfagia.tex}
\newpage

% ------------------------------
% CHAPTER: ECKTORIA
% ------------------------------
\chapter{Ecktoria}
\emph{Marble \& Fire.}

\noindent
{\Large
\textit{My grandfather was a senator. He voted to burn the Ykrul tribute-pact, and then he voted to burn the inquiry into why the grain ships sank. He signed his name in wax and watched the city burn three times. He never raised his voice.}

\vspace{0.5em}

I am a Coin-House Factor. I seal deals with three ribbons—white for "ask," red for "now," black for "never." I have never used the black ribbon. Not because I am kind, but because "never" is a debt I cannot collect.

\vspace{0.5em}

Ecktoria is an Empire that never quite died—more like a corpse propped up by gilded scaffolding and the stubborn refusal of its undertakers to admit the job's done. The Everflame burns in every basilica not as a symbol of divine truth, but as a literal lie detector: swear a false oath near it, and your tongue turns black. I have seen a merchant's tongue turn black mid-sentence. He finished his lie anyway. The buyers pretended not to notice. \textit{That is Ecktoria.}

\vspace{0.5em}

The statues watch. The fire judges. The marble remembers every name carved into its steps—and every name left out.

\vspace{0.5em}

Do not refuse a toast in a coin-house. The ledgers are the true judges, and a declined cup is a declined credit. You will not borrow again.

\vspace{1em}

\noindent\textit{— Marcus of the Red Ledger, Coin-House Factor, third generation}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Spiced flatbread dipped in olive oil and honey—but only if you can afford the honey. The poor eat millet porridge with a single dried fig. The fig is counted as part of the household's grain allowance.

\textbf{What does a child learn before they can walk?} To count coins by touch. Every Ecktorian child has a box of worn sesterces; by age five, they can tell a debased copper from pure silver by the weight alone.

\textbf{What is the first thing you notice about a stranger?} Their signature. Ecktorians judge by handwriting before face. A clerk with a steady hand is trustworthy. A noble with a flourish is hiding something. A merchant who signs with an X is not worth doing business with.

\textbf{What is the most common crime?} Tax evasion—but it is not a crime, it is a sport. Every household keeps two ledgers: the white one for the Censor, the red one for the family. The art is in keeping the red ledger from looking like a confession.

\textbf{What does no one talk about but everyone knows?} The Censor's clerks are paid to find discrepancies. They are also paid to lose them. The price of a lost discrepancy is exactly 5\% of the tax owed. This is not corruption. It is "administrative flexibility."

\textbf{Where do people go when they die?} The family crypt, beneath the basilica floor. The wealthy are burned, their ashes mixed with marble dust and pressed into the floor tiles. The poor are buried in unmarked graves outside the walls. The Everflame's light does not reach that far.

\textbf{How do you apologize for a serious wrong?} You petition the Censor to have the wrong stricken from the public record. The petition costs a tenth of your annual income. If accepted, the wrong is erased—legally, it never happened. If rejected, the wrong is published, and you lose your civic standing.

\textbf{What is the most important festival?} The Triumph of the Flame (autumn equinox). The basilica's central flame is carried in procession through every district. To touch the flame is to be absolved of petty debts. To be burned by it is a sign that your debts are not petty.

\textbf{What is the secret that could destroy a powerful person?} The Grand Magistrate's seal is a forgery. The original cracked in 783 AR. The current seal was carved by a prison engraver who was executed to keep the secret. The engraver's apprentice is still alive. They live in Silkstrand. They have the original sketches.
\end{almanac}

\newpage
\input{../generators/ecktoria.tex}
\newpage

% ------------------------------
% CHAPTER: ACASIA
% ------------------------------
\chapter{Acasia}
\emph{The Broken Marches.}

\noindent
{\Large
\textit{We do not speak of the Curse. We speak around it, under it, through the gaps where words used to fit. The Curse is not a thing. It is the space where a thing used to be.}

\vspace{0.5em}

I am a charcoal-burner. My family has worked these woods for three generations. My grandfather could name every tree between the Pale Causeway and the Wolfstairs Pass. My father could read a deer's track in packed clay. I can tell you where the villages used to be. The map in my head has more empty spaces than marks.

\vspace{0.5em}

Acasia was not always like this. Once, the granaries held grain through winter. Once, the roads carried more than ghosts and tax collectors. Then the trees started dying. Not all at once—slowly, branch by branch, forest by forest. The Blackwood turned grey, then brittle, then silent. The farmers said it was a curse. The duke said it was blight. The merchants said it was the tariffs. The priests said it was the will of gods who had stopped listening.

\vspace{0.5em}

Nobody said it was guilt. Nobody had to. Guilt is the smell of mildew on old parchment. Guilt is the way a village watches you pass without opening its gates. Guilt is the empty chair at the feast, the unmarked grave behind the chapel, the child who will not meet your eyes because their parents told them not to speak to strangers.

\vspace{0.5em}

The truth—the one we do not speak—is that we could have saved the Blackwood. We saw the first trees wither. We saw the sap run black. We did nothing. We told ourselves it would pass. We told ourselves the duke would fix it. We told ourselves the forest was not our responsibility. And then the forest died, and the silence that followed was our answer.

\vspace{0.5em}

The Curse of Acasia is not a spell. It is a memory. The land remembers what we did—and what we failed to do. Every broken milestone on the old Imperial Road is a forgotten promise. Every abandoned vineyard is a harvest we chose not to plant. Every bridge that leads back to the same shore is the road we refused to take.

\vspace{0.5em}

Never sign a contract on a bridge. The river is a witness, and it never forgets a signature. But more than that: the bridge knows when you are lying. Not to the other party. To yourself.

\vspace{0.5em}

We are the Curse now. We carry it in our silence, in our empty fields, in the way we turn away when someone asks where the children went. Acasia is not a broken province. It is a province that broke itself, one small refusal at a time, until there was nothing left to break but the people who stayed.

\vspace{0.5em}

And still we stay. Because leaving would mean admitting we were wrong. And that is the one thing guilt cannot bear.

\vspace{1em}

\noindent\textit{— Corvin, charcoal-burner of the Blackwood, who no longer names the trees because there are too few left to name}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Hardtack soaked in broth—if you have broth. If not, hardtack and hope. The granaries have been empty for three winters. The merchants sell grain at famine prices. The duke says the roads are too dangerous for caravans. The farmers say the duke owns the caravans.

\textbf{What does a child learn before they can walk?} To name the dead. Every Acasian child has a list of ancestors they repeat at bedtime. The list gets shorter each generation. Not because children forget. Because the dead are forgetting themselves.

\textbf{What is the first thing you notice about a stranger?} Whether they have a shadow. The Curse takes shadows first. A man without a shadow is not cursed—he is already gone. His body just hasn't stopped moving yet.

\textbf{What is the most common crime?} Bridge-toll evasion. The roads are unsafe, the ferries are unreliable, but the barons still demand payment for crossings no one can make. The crime is not in refusing to pay. The crime is in being caught trying.

\textbf{What does no one talk about but everyone knows?} The Curse began with a promise broken on the Blackwood Bridge. A baron promised his daughter to a lord of the old forest. He married her to a lowland knight instead. The forest remembered. The forest is still waiting for the wedding.

\textbf{Where do people go when they die?} Into the Blackwood. Burial is forbidden—the ground rejects the dead. Bodies are carried to the tree-line and left. The wood takes them. Sometimes, at dusk, you can see them walking between the trunks. They do not look back.

\textbf{How do you apologize for a serious wrong?} You walk the Broken Mile—a stretch of road that collapsed into the river a decade ago. If you reach the far side, you are forgiven. If you fall, you were not sorry enough. No one has reached the far side in seven years.

\textbf{What is the most important festival?} The Counting of the Names (midwinter). Every village lists its dead aloud, from oldest to newest. If a name is forgotten, the ghost walks until someone remembers. The list grows longer each year. The memory grows shorter.

\textbf{What is the secret that could destroy a powerful person?} The duke's son is a ghoul. He was lost in the Blackwood as a child. They found him three days later, pale and smiling. He does not age. He does not eat. He only watches. The servants know. The servants do not speak.
\end{almanac}

\newpage
\input{../generators/acasia.tex}
\newpage

% ------------------------------
% CHAPTER: VHASIA
% ------------------------------
\chapter{Vhasia}
\emph{The Fractured Sun.}

\noindent
{\Large
\textit{Two crowns. One kingdom. A river so wide it might as well be a border—and oh, how the crowns love pretending it's not.}

\vspace{0.5em}

I am a Road Warden. I have served under three different banners on this same stretch of road, and the only constant is that someone always wants to collect tolls. The Northern Seigneurie and the Southern Parlement both claim the throne, both flood the river with paperwork to "prove" their right, and both are too busy arguing over seal wax colours to notice that the Fenwoods already burned Lence, and the Sun-court is a memory.

\vspace{0.5em}

The Last Dauphin? A rumour. But rumours pay well when you're the one who started them—and I started three: that he was swapped at birth with a goatherd's child, that he's secretly a Veiled Sister in disguise, and that he's been dead for ten years, his "reappearance" a puppet show run by the Seigneurie's spymaster.

\vspace{0.5em}

Never carry a coin that bears only one sunburst. The realm has three mints, and the wrong metal on the wrong road is treason.

\vspace{1em}

\noindent\textit{— Road Warden Cerran, who has not slept in two days and is not sure which crown he serves}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Dark bread and watered wine—the wine cuts the dust from the road. The nobility eat spiced eggs and fresh cheese, but they eat behind closed doors. To eat well in public is to invite envy. Envy is a dagger.

\textbf{What does a child learn before they can walk?} Which colours not to wear. Black and gold are the Northern Crown's colours. Red and silver are the Southern Parlement's. A child who wears the wrong colour to the wrong village may not return home.

\textbf{What is the first thing you notice about a stranger?} Which side of the road they walk on. The north side is the Crown's. The south side is the Parlement's. The middle is no man's land—literally. Bandits, spies, and messengers use the middle. So do ghosts.

\textbf{What is the most common crime?} Raiding across the Belworth. The river is the border, the border is a suggestion, and the suggestion is enforced by armed men on both sides. The crime is not in raiding. The crime is in being caught on the wrong bank at dawn.

\textbf{What does no one talk about but everyone knows?} The Sun-court was not burned. It was bankrupted. The Fenwoods did not sack Lence—they bought it. Every judge, every clerk, every guard captain had a price. The Fenwoods paid it. The Vhasians are still paying interest.

\textbf{Where do people go when they die?} To the crypts beneath the Sun-court—if they can afford a plaque. The poor are buried in unmarked graves along the Belworth floodplain. The river floods every spring. The dead wash downstream, into the land of the other crown. Neither side claims them.

\textbf{How do you apologize for a serious wrong?} You offer a chevauchée—a raiding party that burns a village in your name. The apology is not in the burning. The apology is in the promise that the village will be rebuilt with your own hands. If you burn a village and do not rebuild, you have not apologized. You have started a feud.

\textbf{What is the most important festival?} The Sun-return (spring equinox). Both crowns claim the sun as their symbol. On this day, both hold processions. The processions are identical. The participants do not acknowledge each other. The sun does not care.

\textbf{What is the secret that could destroy a powerful person?} The Queen-Mother has no claim to the regency. Her son is not the Dauphin. The real Dauphin died in childbirth. The Queen-Mother's daughter was stillborn. The child on the throne is a foundling, bought from a Tulkani caravan. The Tulkani know. The Tulkani are still waiting to be paid.
\end{almanac}

\newpage
\input{../generators/vhasia.tex}
\newpage

% ------------------------------
% CHAPTER: VITERRA
% ------------------------------
\chapter{Viterra}
\emph{The Hedge-Law Realm.}

\noindent
{\Large
\textit{Every hedge here marks not just a boundary but a principle—the careful balance between order and freedom, justice and mercy. Our law grows like the hedges themselves, tended with precision and patience.}

\vspace{0.5em}

I am a Queen's Justiciar. I have never lost a legal duel. My opponents have a habit of falling down stairs. I do not push them. The stairs \textit{remember their debts.}

\vspace{0.5em}

Viterra breathes with the careful rhythm of precedent and petition. Once a proud kingdom before becoming a Utaran province, it is now reborn under a Warrior Queen who rules by legal precision as much as military might. The realm's true sinews lie not in steel but in writs, charters, and the ancient rights that bind lord to tenant, crown to county, and church to common.

\vspace{0.5em}

But do not mistake us for gentle. The Queen's Peace is not a promise—it is a contract, renewed each morning by every magistrate, every justicar, every clerk who opens a ledger. And across the Belworth, the Vhasians still claim our hedges are theirs. Let them try.

\vspace{0.5em}

Never cross the Belworth without a charter recognized on both banks. Viterran law and Vhasian claim fight every crossing, and the river takes the losers.

\vspace{1em}

\noindent\textit{— Justiciar Elodie Venn, who invoked the Right of Acknowledgment and watched a father speak truth to his son for the first time}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Barley porridge with a drizzle of honey—if you own bees. If not, porridge with salt. Salt is the currency of the poor. Honey is the currency of the hopeful.

\textbf{What does a child learn before they can walk?} To recite their lineage back three generations. Every Viterran child knows their grandparents' names, their grandparents' trades, and their grandparents' debts. A child who cannot recite their lineage is a child without standing. Standing is everything.

\textbf{What is the first thing you notice about a stranger?} Whether they carry a seal ring. A seal ring means they have the right to make contracts in their own name. A stranger without a seal ring is either a servant or a criminal. In Viterra, the difference is a matter of paperwork.

\textbf{What is the most common crime?} Boundary disputes. A hedge moved three feet east. A milestone replaced under cover of darkness. A ford declared "private" when it has been public for two centuries. The courts are clogged with such cases. The lawyers are rich. The hedges are patient.

\textbf{What does no one talk about but everyone knows?} The Queen is not a warrior. She has never drawn a sword. Her reputation was forged by a propaganda campaign run by her spymaster. The spymaster is dead. His successor is more discreet.

\textbf{Where do people go when they die?} To the family plot—if they have a family. The landless poor are buried in parish cemeteries, their headstones marked with a single word: "Anonymous." The ghosts of the anonymous wander the hedgerows at dusk. They do not remember their names. They remember their debts.

\textbf{How do you apologize for a serious wrong?} You petition the hedge-court for a "writ of contrition." The writ requires you to state your wrong aloud, in public, while holding a handful of earth from the wronged party's land. If the earth crumbles, you are forgiven. If it holds, your apology is not accepted. You may try again next season.

\textbf{What is the most important festival?} The Opening of the Rolls (autumn). Every county sends its tax records to the capital. The procession is a week long. The documents are carried by children, so that the next generation may learn the weight of obligation.

\textbf{What is the secret that could destroy a powerful person?} The Codex of Precedents has a missing page. It was torn out fifty years ago, during a trial for treason. The page contained the precedent for executing a monarch. The Queen knows. The Justiciars know. The page is hidden in a lead-lined box beneath the hedge-court at Thornwood. The hedge-court is haunted. The ghost knows where the key is.
\end{almanac}

\newpage
\input{../generators/viterra.tex}
\newpage

% ------------------------------
% CHAPTER: THEPYRGOS
% ------------------------------
\chapter{Thepyrgos}
\emph{City of a Thousand Stairs.}

\noindent
{\Large
\textit{Every stair climbed is a word in the great conversation between earth and sky. This city is not built—it is grown, stone by stone, law by law, until the very architecture speaks wisdom.}

\vspace{0.5em}

I am a Bell-Runner. I know every step that creaks a warning and every landing that whispers gossip. The bells tell the city's mood better than any lord—three chimes means business, six means trouble, and nine means the hunters are coming.

\vspace{0.5em}

The Synod rules here with the delicacy of a brick wrapped in velvet—they hate unsanctioned magic like it's a personal insult, but they adore infrastructure, especially when it's shiny, expensive, and comes with a side of indulgences they can sell to the faithful. The Chain-Lanterns move at dawn, their brass censers swinging, their bells tuned to the frequency of latent magic. In the university's forbidden tiers, the Witness and the Ninth arm rival fraternities with citations and curses, and no one is sure which side the Archon secretly favours.

\vspace{0.5em}

Never let a bell-ringer count your breaths. The Chain-Lanterns pay well for names that ring false on the test stairs.

\vspace{1em}

\noindent\textit{— Sofia, Bell-Runner, who ties a knot in her hair for each hunter she has outrun}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Flatbread with olive oil and a boiled egg—if you are a clerk. If you are a student, bread and water. If you are a professor, bread, water, and a single fig. The fig is a status symbol. The status is "I have not been audited."

\textbf{What does a child learn before they can walk?} To read a bell schedule. Thepyrgos runs on bells—dawn bells, market bells, curfew bells, trial bells. A child who cannot read the bells cannot cross the city without a guardian. A child without a guardian is a child who will be sold to the university for dissection.

\textbf{What is the first thing you notice about a stranger?} Their step count. Thepyrgos has a thousand stairs. A stranger who counts them aloud is either a pilgrim or a spy. A stranger who climbs them in silence is either a native or a ghost.

\textbf{What is the most common crime?} Unsanctioned magic. The Synod licenses every spell cast within the walls. An unlicensed cantrip is a misdemeanour. An unlicensed healing is a felony. An unlicensed summoning is a death sentence—not for the caster, who will be executed, but for the summoned, who will be bound to the city's foundation as a "structural consultant."

\textbf{What does no one talk about but everyone knows?} The Silent Stair is not silent. It whispers. The whispers are the confessions of the dead—scholars who died in the university's service, their secrets recorded in the stone. The Chain-Lanterns listen. They do not share.

\textbf{Where do people go when they die?} To the catacombs beneath the university. The catacombs are a library of bones. Each skeleton is labelled with the scholar's name, their thesis, and the date of their death. The unlabelled bones are the ones who died before they could finish. They are still writing.

\textbf{How do you apologize for a serious wrong?} You submit a "retraction" to the Synod's Committee on Ethical Correction. The retraction must be published in the public square, read aloud by a herald, and witnessed by at least three notaries. The fee is 50 silver. The shame is free.

\textbf{What is the most important festival?} The Convocation of the Steps (summer solstice). Every citizen climbs the thousand stairs to the upper citadel. At the top, the Archon reads the names of the dead from the past year. The names are carved into the steps as they are read. The steps are getting crowded.

\textbf{What is the secret that could destroy a powerful person?} The Archon is a fraud. He never passed the Synod's examination. His degree was purchased from a disgraced dean whose signature is forged on every document. The forger is still alive. They live in the catacombs. They have the original quill.
\end{almanac}

\newpage
\input{../generators/thepyrgos.tex}
\newpage

% ============================================================
% PART II: THE NORTHERN REACHES
% ============================================================
\part{The Northern Reaches}
% After \part{The Northern Reaches}
\begin{quote}
\small
\textit{``The mountain counts your breaths. The frost writes your name in ice. In the north, the dead do not wait – they walk, and they remember every promise you made to keep warm.''}
\noindent — Dusana of the Raven Road
\end{quote}
% ------------------------------
% CHAPTER: AELER
% ------------------------------
\chapter{Aeler}
\emph{Crowns \& Under-Vaults.}

\noindent
{\Large
\textit{The mountain remembers every word carved in its stone, every seal pressed into its walls. To govern here is not to command, but to harmonize with the ancient songs that still echo in the deep places.}

\vspace{0.5em}

I am an Under-Mason. I can hear stone shift a room away. When the rock starts singing a different tune, you best listen. Three knocks in the morning means all's well, but three knocks at night means something's coming through the walls that wasn't meant to walk.

\vspace{0.5em}

My homeland. The place where we don't trust the surface, but we do trust the weight of a well-placed keystone—and oh, how the surface loves to test that trust. Our breath-bonds aren't just currency—they're the hardest thing in the peaks, forged in the sighs of miners who knew too much. We sell them to Vilikari jarls who pay in furs stained with brother's blood, to Ykrul khagans who weigh the value in warhorses and broken oaths. Once, we sold one to the Veiled Sisters. They paid in secrets—which is another way of saying they paid in future leverage.

\vspace{0.5em}

Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one—a breath, a memory, a shadow. My grandfather reached nine once. He lived another thirty years, but he never cast a shadow again.

\vspace{1em}

\noindent\textit{— Thromm Stone-Tongue, Under-Mason, who has never counted to nine}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Stone-bread and cave cheese—the bread is baked on heated stone slabs, the cheese is aged in sealed galleries where the temperature never varies. The cheese has been aging for three generations. It is an heirloom. It is also barely edible.

\textbf{What does a child learn before they can walk?} To count their breaths. Aeler children wear breath-counters—small brass bells on leather cords. Eight breaths, ring the bell. Nine breaths, run. A child who cannot count is a child who will not survive the winter.

\textbf{What is the first thing you notice about a stranger?} Their bell. Every Aeler carries a bell tuned to their breath. A stranger's bell will ring at a different pitch—the mountain's lungs are calibrated to local frequencies. A mismatched bell is a warning. A silent bell is a threat.

\textbf{What is the most common crime?} Air theft. In the deep vaults, ventilation is shared. A sealed door, a blocked vent, a diverted shaft—these are thefts of breath, punishable by breath-debt. The guilty party must donate their own air ration to the victim. The donation is measured in days. The days are counted by bells.

\textbf{What does no one talk about but everyone knows?} The True Masons are not neutral. They are the Vault-Queen's eyes. Every "independent" repair, every "neutral" inspection, every "routine" audit—the True Masons are watching. They do not report to the Queen. They are the Queen. They have always been the Queen.

\textbf{Where do people go when they die?} To the Deep Vaults—the oldest galleries, where the air is thin and the stone is cold. The dead are not buried. They are sealed in niches carved into the walls, their names chiselled above their heads. The niches are reused every century. The names are never erased. The dead share walls with strangers. The strangers are patient.

\textbf{How do you apologize for a serious wrong?} You offer a keystone—a load-bearing stone from your own holding. The stone is installed in the wronged party's wall. If it holds, you are forgiven. If it cracks, the debt stands. If it crumbles, your holding collapses. Apology is architecture. Architecture is law.

\textbf{What is the most important festival?} The Ringing of the Underways (deep winter). Every bell in every hold is rung at the same hour—eight times, no more. The sound travels through the stone, a heartbeat that reminds the mountain that the living are still here. The dead listen. The dead do not ring back.

\textbf{What is the secret that could destroy a powerful person?} The Vault-Queen is a construct. She was assembled in 723 AR, after the real queen died in a cave-in. Her heart is a keystone. Her breath is a bellows. Her voice is a recording. The True Masons maintain her. They have the schematics. The schematics are missing three pages. The missing pages describe how to turn her off.
\end{almanac}

\newpage
\input{../generators/aeler.tex}
\newpage

% ------------------------------
% CHAPTER: THE MISTLANDS
% ------------------------------
\chapter{The Mistlands}
\emph{Bells, Salt, and Breath.}

\noindent
{\Large
\textit{The bells here do more than mark time—they weave the very fabric of reality, keeping the mist at bay and the dead in their proper places. Each chime is a thread in the pattern that holds the world together.}

\vspace{0.5em}

I am a Bell-Warden. I refuse to write down the ninth note. It lives only in my memory—and that memory is fading.

\vspace{0.5em}

The Mistlands are a protectorate of the Aeler in name only. The fog here doesn't just hide the land; it hides the law, and oh, how the wraiths love that ambiguity. The bell-lines aren't just defences—they're the only thing keeping the wraiths from dragging whole villages into the mist. The Aeler built them. The Aeler charge for them. The Aeler will leave when the money runs out. And when that happens, the fog will remember every single soul they abandoned.

\vspace{0.5em}

Never answer a voice that calls your name from the fog. The mist collects names, and once it has yours, the ferryman will always know where to find you. I've seen men answer. I've watched them walk into the white, still talking, still smiling, until their voices came back from a direction they hadn't gone.

\vspace{1em}

\noindent\textit{— Warden Kael, who has seen the ninth bell rung once and will not speak of it}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salt fish and sour bread—the fish is smoked over peat fires, the bread is leavened with fog-damp yeast. The yeast is caught in cloth traps hung from the bell-towers. The traps are emptied at dawn. The yeast is not alive. It is not dead either.

\textbf{What does a child learn before they can walk?} To ring a bell. Every Mistlands child is given a small bronze bell at birth. By age three, they can ring the eight notes of the ward-chime in their sleep. The ninth note is never taught. The ninth note is discovered. Discovery is fatal.

\textbf{What is the first thing you notice about a stranger?} Their salt-pouch. Every Mistlander carries a leather pouch of consecrated salt. A stranger without a pouch is either a fool or a wraith. A stranger whose pouch is empty is a fool who is about to become a wraith.

\textbf{What is the most common crime?} Ward-salt theft. The Aeler control the salt pans, and the salt is the only thing that keeps the dead from rising. A stolen pouch can buy a week's protection. The price is a year's debt to the Aeler. The debt is collected in labour. The labour is collecting salt. The salt is stolen. The cycle is perpetual.

\textbf{What does no one talk about but everyone knows?} The Bell-Wardens are not chosen—they are volunteered. A village that loses a warden must send a replacement to the Aeler training holds. The holds are in the mountains. The mountains are cold. The trainees do not always return. The ones who do return do not speak of what they learned. Their bells ring at a different pitch.

\textbf{Where do people go when they die?} To the fog. Bodies are not buried in the Mistlands—the ground rejects them. Corpses are carried to the edge of the bell-line and left for the mist. The mist takes them. Sometimes, at dusk, you can see them walking the lanes. They do not speak. They do not need to. Their presence is a question.

\textbf{How do you apologize for a serious wrong?} You ring the Bell of Atonement—a cracked bronze bell that hangs in every village square. The bell rings once for each wrong confessed. If the bell rings true, you are forgiven. If it cracks further, you are not. If it shatters, you are banished to the fog. The banished never return. Their voices still answer from the white.

\textbf{What is the most important festival?} The Rehanging of the Lines (autumn). Every bell-line is taken down, cleaned, and rehung. The work takes three days. The villages are silent for three nights. The dead listen. The dead wait. The dead do not ring.

\textbf{What is the secret that could destroy a powerful person?} The Legate of the Mists is a wraith. She died in 803 AR, during a bell-line failure. The Aeler rebuilt her from fog and memory. She does not know she is dead. She attends council meetings. She signs documents. Her signature fades after sunset. The documents are refiled every morning.
\end{almanac}

\newpage
\input{../generators/mistlands.tex}
\newpage

% ------------------------------
% CHAPTER: YKRUL
% ------------------------------
\chapter{Ykrul}
\emph{Wolf Standards, Winter Camps.}

\noindent
{\Large
\textit{The steppe remembers every oath sworn beneath its endless sky, and the wolf-standards carry the honour of clans that have outlasted empires. To lead here is to keep the old ways sharp as winter steel.}

\vspace{0.5em}

I am a Herd-Scout. I can read a trail by the way the grass bends—and a lie by the way a rider sits. The Ykrul don't measure wealth in silver; we measure it in hoofbeats, in the snort of a warhorse at dawn, in the weight of a promise kept through winter.

\vspace{0.5em}

We do not build monuments. We build relationships, each one a stone in an invisible cairn that stretches from the Dolmis to the Aberderrin. A camp is a weave of favours owed and honour satisfied. A feud is a slow equation: blood in, blood out, with interest compounded by winter.

\vspace{0.5em}

The Khagan's Court doesn't rule from a throne—it rules from a saddle. Our "laws" are the Thirty-Six Sayings of the Road, and the thirty-seventh is the one you learn alone, on a night when the stars blink out and the wolf does not howl.

\vspace{0.5em}

Never ride between two wolves at dusk. The space between them is a blind angle—a route with no exit, a debt with no payment. The wolves know this. They choose the gap deliberately.

\vspace{1em}

\noindent\textit{— Sogata, Herd-Scout, who has never been lost but has been found many times}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Fermented mare's milk and dried meat—the milk is drunk warm from the skin, the meat is sliced thin and dried on saddle-pommels. The meat carries the taste of horse sweat. The milk carries the taste of the steppe. Both are acquired. Both are necessary.

\textbf{What does a child learn before they can walk?} To sit a horse. Ykrul children are placed on a horse before they can stand. The horse is old, patient, and knows the child's weight. A child who falls is a child who learns to hold tighter. A child who cannot ride is a child who cannot eat.

\textbf{What is the first thing you notice about a stranger?} Their braids. Ykrul warriors braid their hair with clan colours—red for the Wolf Clan, black for the Iron Horse, grey for the Salt Steppe. A stranger whose braids are cut has been shamed. A stranger whose braids are dyed black has killed a clan-mate. A stranger with no braids has no clan. A stranger with no clan is prey.

\textbf{What is the most common crime?} Horse theft. A stolen horse is a stolen relative. The price is blood—the thief's or the owner's. The courts are the council fires. The evidence is the horse's memory. Horses remember faces. Horses remember smells. Horses do not lie.

\textbf{What does no one talk about but everyone knows?} The Khagan is not chosen by blood. He is chosen by the horse. Every five years, the clan chiefs lead their best stallions to the sacred ring. The horse that steps forward first chooses the next Khagan. The horse has not stepped forward in twenty years. The council chooses by silence. The silence is growing loud.

\textbf{Where do people go when they die?} To the sky. Ykrul dead are placed on open platforms, wrapped in felt, and left for the wind and the birds. The bones are collected after a year and ground into powder. The powder is mixed with clay and fired into beads. The beads are worn by the dead's descendants. The beads are warm. The warmth is memory.

\textbf{How do you apologize for a serious wrong?} You offer a hostage string—a braided cord that ties your wrist to the wronged party's. The cord is cut when the apology is accepted. If the cord is cut before the apology is finished, you are not forgiven. If the cord is never cut, you are bound. The binding is lifelong. The binding is called "friend."

\textbf{What is the most important festival?} The First Grass (spring equinox). The herds are released onto the new pasture. The riders race to the horizon and back. The winner names the summer camps. The loser tends the sick animals. The sick animals are always the loser's. The loser always wins next year.

\textbf{What is the secret that could destroy a powerful person?} The Sky-Speaker is a fraud. She cannot call storms. She cannot read omens. She cannot speak to the dead. She is a horse-witch who learned to read the clouds. The clouds are reliable. The Khagan does not know. The Khagan does not want to know.
\end{almanac}

\newpage
\input{../generators/ykrul.tex}
\newpage

% ------------------------------
% CHAPTER: VILIKARI
% ------------------------------
\chapter{Vilikari}
\emph{Laurels \& Longhouses.}

\noindent
{\Large
\textit{We are the bridge between empire and wilderness, the marriage of law and custom. Our strength lies not in choosing one path but in walking both with equal surety.}

\vspace{0.5em}

I am a Hearth-Mother. I have never lost a quarrel. My voice is soft, my memory longer than the Foedus Stone.

\vspace{0.5em}

The Vilikari straddle the border like a drunkard on a fence—half empire, half steppe, all headache. Their foedus treaties aren't just agreements—they're a tragicomic mess, half carved in marble by emperors who forgot to pay the sculptors, half braided in string by chieftains who forgot which knot meant "peace." And yet, somehow, the treaties hold. Not because of the marble. Not because of the string. Because both sides are too exhausted to start the argument over again.

\vspace{0.5em}

Never invoke only one law. The mallus (oak-ring) and the basilica (marble bench) must both be honoured. Choose one, and the other will find you—usually at midnight, usually with a writ, usually with a sword.

\vspace{1em}

\noindent\textit{— Hearth-Mother Vesna, who has fed three kings and forgotten none of their debts}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Oatmeal with dried berries—the berries are harvested from the border scrub, where the empire's orchards meet the steppe's wild brambles. The berries taste of both. The oatmeal tastes of neither. The oatmeal is filler. The berries are the point.

\textbf{What does a child learn before they can walk?} To speak both languages—Imperial Utaran and the steppe cant. A child who speaks only one is a child who will never cross the border. A child who crosses the border is a child who will come back with stories. A child with stories is a child who will be listened to. Listening is power.

\textbf{What is the first thing you notice about a stranger?} Which side of the road they ride on. The empire's roads are measured from the centre. The steppe's paths are measured from the edge. A stranger who rides in the centre is an imperial. A stranger who rides at the edge is a nomad. A stranger who rides in the ditch is a spy. The ditch is not a road. The ditch is a message.

\textbf{What is the most common crime?} Treaty-splitting—selling the same rights to both sides. A Vilikari factor will promise river access to an imperial merchant and grazing rights to a Ykrul clan, then collect from both and disappear before the paperwork collides. The crime is not in the selling. The crime is in getting caught before you cross the border.

\textbf{What does no one talk about but everyone knows?} The Foedus Stone is a copy. The original was stolen in 712 AR by a Vilikari trader who owed debts on both sides. She melted it down and sold the marble dust as "treaty-strengthening powder." The powder was mostly chalk. The treaties are still holding. The traders are still selling.

\textbf{Where do people go when they die?} To the border. The Vilikari bury their dead at the edge—close enough to the empire for the priests to pray, close enough to the steppe for the shamans to chant. The dead are planted face-up, so they can see both skies. The skies are different colours. The dead do not complain.

\textbf{How do you apologize for a serious wrong?} You invoke the mallus—the oak-ring court. The wronged party sits inside the ring. The wrongdoer circles it three times, speaking their apology. On the third circuit, the elders judge. If the apology is accepted, the wrongdoer is admitted into the ring. If it is rejected, the ring closes. The wrongdoer is outside. Outside is the lawless land. The lawless land is where the wolves are.

\textbf{What is the most important festival?} The Renewal of the Foedus (summer solstice). The stone is washed with milk and honey. The milk is from a Ykrul mare. The honey is from an imperial apiary. The mixture is drunk by the elders. The elders speak the treaty aloud. The treaty is long. The elders are old. The ceremony takes all day.

\textbf{What is the secret that could destroy a powerful person?} The Foedus Stone has a crack. The crack was caused by a hammer in 799 AR, during a dispute over horse tariffs. The stone was repaired with clay and prayer. The clay is drying. The prayer is fading. The crack is spreading. When the stone splits, the treaty splits with it. No one knows what happens then. No one wants to find out.
\end{almanac}

\newpage
\input{../generators/vilikari.tex}
\newpage

% ------------------------------
% CHAPTER: BLACK BANNERS
% ------------------------------
\chapter{Black Banners}
\emph{Condotta \& Crowns.}

\noindent
{\Large
\textit{A contract is only as good as the next payday, and a grudge is only as sharp as the sword backing it. In this game, we sell steel by the season and vendettas by the year.}

\vspace{0.5em}

I am a Veteran Sergeant. I've seen three captains rise and fall in the time it takes to grow a beard, but the mud stays the same and the pay comes when it pleases.

\vspace{0.5em}

Mercenary companies sell steel to the highest bidder—and oh, how they love a good spreadsheet. The Banner-Captains don't just want payment; they want proof they're getting their money's worth, which is why they pay in safe-conducts that are good until the next battle... and utterly worthless after, once the smoke clears and the ledgers need balancing.

\vspace{0.5em}

The camp sings when payday comes. The camp falls silent when the ledger is written in blood. I've heard both. I prefer the singing. At least the singing means you're still alive.

\vspace{0.5em}

Never break a Bell Dawn. Three bells at sunrise means bury the dead, trade prisoners, fix the wells, no ambushes. Violating it is how legends end badly—and how captains end hanging.

\vspace{1em}

\noindent\textit{— Old Sorka, Veteran Sergeant, who counts scars like coins and spends them on whiskey and mercy}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Hardtack and camp stew—the stew is whatever was caught, killed, or traded for before dawn. The cook is called "The Alchemist." No one knows their real name. No one asks.

\textbf{What does a child learn before they can walk?} To count coin. Camp children are taught to tally pay shares before they can read. A child who miscounts is a child who will go hungry. A child who goes hungry is a child who learns to count faster.

\textbf{What is the first thing you notice about a stranger?} Their colours. Every Black Banner company has its own banner—lantern, thorn, bell, fishbone, broken scale. A stranger whose colours are faded has been running. A stranger whose colours are too bright has never fought. A stranger with no colours has no company. A stranger with no company is a deserter. A deserter is a liability. A liability is a corpse.

\textbf{What is the most common crime?} Desertion—but desertion is not a crime, it is a negotiation. A soldier who wants to leave will sell their gear to the quartermaster, sell their secrets to the enemy, and sell their silence to the captain. The price is negotiable. The price is always "one more season."

\textbf{What does no one talk about but everyone knows?} The Black Colonel is not a person. The Colonel is a title held by a wax seal. The seal is kept in a lead-lined box. The box is carried by a mute lieutenant. The lieutenant signs documents with the Colonel's thumb-print. The thumb-print is a forgery. The forger is the lieutenant. The lieutenant has no thumbs.

\textbf{Where do people go when they die?} To the pyre. Camp dead are burned in a bonfire of broken spears and banner poles. The ashes are scattered by the quartermaster. The quartermaster charges a fee. The fee is deducted from the dead's final pay. The final pay is never enough.

\textbf{How do you apologize for a serious wrong?} You stand a watch. The wronged party chooses the hour—midnight, dawn, the dead of winter. You stand alone. If you survive, you are forgiven. If you do not, the debt is paid. The camp does not mourn. The camp does not forget.

\textbf{What is the most important festival?} Payday—but payday is not a day, it is a negotiation. The captain promises payment on the first of the month. The quartermaster promises payment when the supply train arrives. The soldiers promise loyalty until the payment comes. The payment never comes. The loyalty never ends.

\textbf{What is the secret that could destroy a powerful person?} The Banner-Captain's contract is written in disappearing ink. The ink was supplied by a Sidhi forger who was paid in promises. The promises were not kept. The forger is now a camp follower. The forger knows where the contract is hidden. The forger is waiting for the right offer.
\end{almanac}

\newpage
\input{../generators/black_banner.tex}
\newpage

% ------------------------------
% CHAPTER: THE LINNS
% ------------------------------
\chapter{The Linns}
\emph{Skerries \& Storm-Oaths.}

\noindent
{\Large
\textit{The sea keeps better counsel than any scribe, and the storm remembers every oath sworn upon its winds. To rule these shores is to dance between their moods and never miss a step.}

\vspace{0.5em}

I am a Net-Wife. I tie a knot for every storm I've survived. My belt has thirty.

\vspace{0.5em}

North of the Aberderrin Sea, the Linns don't just raid and trade—they do it with the fervour of the converted, swearing oaths to storm and sea that aren't just words—they're leverage with a supernatural clause. They pay in stories, not silver, because to them, a tale well-told is worth more than a chest of gold... especially when it's true.

\vspace{0.5em}

The jarls measure their wealth in ships. The skalds measure their wealth in verses. The dead measure their wealth in how many living remember their names. I measure mine in knots. Each knot is a promise I kept to the sea. The sea remembers. The sea does not forgive forgetting.

\vspace{0.5em}

Never name a drowned person aloud while underway. The sea collects names, and if it hears you speaking one it doesn't recognize, it may reach up to correct the ledger.

\vspace{1em}

\noindent\textit{— Astrid, Net-Wife, who has never lost a sailor to the mist but has lost three to bad stories}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Seal blubber and flatbread—the blubber is chewy, the bread is dense, and both are stored in the same barrel. The barrel is shared with the crew. The crew shares everything. The crew is family. Family is obligation.

\textbf{What does a child learn before they can walk?} To tie a knot. Linn children are given a length of rope at birth. By age five, they can tie the eight sea-knots: clew, clove, figure-eight, reef, sheet, stopper, timber, and the ninth—the one that slips when pulled. The ninth knot is for emergencies. The ninth knot is for funerals. The ninth knot is for when the sea asks for a life.

\textbf{What is the first thing you notice about a stranger?} Their breath. Linn sailors chew kelp to sweeten their breath—the kelp is harvested from sacred beds, blessed by the Volva. A stranger whose breath smells of salt has been at sea. A stranger whose breath smells of nothing has not. A stranger whose breath smells of copper has spoken a false oath. The sea knows. The sea will answer.

\textbf{What is the most common crime?} Oath-breaking—but oath-breaking is not a crime, it is a storm. The oath-breaker's ship will founder. The oath-breaker's nets will tear. The oath-breaker's children will be born with webbed fingers. The punishment is not imposed by the jarl. The punishment is imposed by the sea. The sea is patient. The sea does not forget.

\textbf{What does no one talk about but everyone knows?} The Sea-Queen is not a queen. She is a figurehead—literally. Her body is carved from the keel of a ship that sank in 612 AR. Her voice is a recording. Her throne is a navigation table. The jarls pretend. The jarls have always pretended. The pretending keeps the peace.

\textbf{Where do people go when they die?} To the waves. Linn dead are placed in boats and set adrift, weighted with stones. The stones are carved with the dead's name. If the boat sinks before it reaches the horizon, the dead is accepted. If it reaches the horizon, the dead is rejected. If it returns to shore, the dead is cursed. The cursed wash up on beaches. The beaches are littered with their bones.

\textbf{How do you apologize for a serious wrong?} You offer a storm-oath. You stand on a cliff during a gale and swear your apology to the wind. If the wind carries your words, you are forgiven. If the wind throws you into the sea, you are not. The sea does not give second chances.

\textbf{What is the most important festival?} The Launching of the Longships (spring thaw). The ships are dragged from their winter shelters and blessed with blood and mead. The blood is from a sacrificed goat. The mead is from the jarl's own stores. The ships are named after ancestors. The ancestors are listening. The ancestors approve.

\textbf{What is the secret that could destroy a powerful person?} The Volva of the Mist is a fraud. She cannot see the future. She reads the weather—the pressure, the wind, the taste of salt on the air. The sailors know. The sailors do not care. A weather-reader is more useful than a seer. The Volva knows this. The Volva uses this. The Volva is never wrong.
\end{almanac}

\newpage
\input{../generators/linns.tex}
\newpage

% ============================================================
% PART III: THE EASTERN DOLMIS & THE VALEWOOD
% ============================================================
\part{The Eastern Dolmis \& The Valewood}

% After \part{The Eastern Dolmis & The Valewood}
\begin{quote}
  \small
  \textit{``Every threshold has a keeper. Some keepers ask for salt. Some ask for silence. The Green Host asks for a name – and it does not mean yours.''}
  \noindent — Dusana of the Raven Road
  \end{quote}
% ------------------------------
% CHAPTER: THEONA
% ------------------------------
\chapter{Theona}
\emph{Three Greens, No Ninth.}

\noindent
{\Large
\textit{The wells here hold more than water—they hold the dreams and prayers of generations. To govern is to tend these sacred spaces where the living commune with what came before.}

\vspace{0.5em}

I am a Peat-Cutter. My barrow has a false bottom. I hide debts there—and bodies.

\vspace{0.5em}

The isles of Theona rise green and jagged from the Dolmis, their basalt cliffs wrapped in mists and moorland song. Nominally a duchy sworn to Viterra, in truth the Three Greens heed no crown but their own moots, no law but their taboos. The greatest of these is the silence of Nine: no bell tolls a ninth, no bridge counts that span, no oath dares its syllable.

\vspace{0.5em}

The Viterran sheriffs call it superstition. The Green Host calls it trespass. The difference is that the sheriffs will hang you. The Green Host will take your name, and you will wander the mist until you forget you ever had one.

\vspace{0.5em}

Never pour the ninth cup at a feast. Never speak the ninth name in a bargain. Never step where a step is missing. The Green Host rides for those who complete what should remain incomplete.

\vspace{1em}

\noindent\textit{— Celia, Hedge-Witch Midwife, who has tied nine thousand knots and untied nine}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Oatcakes and heather ale—the oats are grown on terraced hillsides, the heather is harvested from the high moors. The ale is brewed in secret, by women who have forgotten more recipes than the men will ever learn. The men do not ask. The men are wise.

\textbf{What does a child learn before they can walk?} To count to eight. Theona children are taught numbers on their fingers—eight fingers, two thumbs. The thumbs are not counted. The thumbs are for pointing. Pointing is for warnings. Warnings are for the ninth thing.

\textbf{What is the first thing you notice about a stranger?} Their shadow. Theona is a land of mists and twilight. A stranger whose shadow is too short has been touched by the Host. A stranger whose shadow is too long is running from them. A stranger with no shadow is already taken. Do not speak to them. Do not meet their eyes. Do not offer them salt.

\textbf{What is the most common crime?} Breaking the ninth taboo—but breaking the taboo is not a crime, it is an invitation. The Host rides for those who pour the ninth cup, speak the ninth name, or take the ninth step. The Host does not arrest. The Host takes. The taken are never found. Their empty cups remain on the table. The cups are warm for three days.

\textbf{What does no one talk about but everyone knows?} The Green Host are not ghosts. They are the living—Theonans who broke the taboo and were taken. They ride with the King in Green, who was once a chieftain of the northern isle. He broke the ninth vow. He was not taken. He was crowned. His crown is made of thorns. His throne is a barrow. His court is the mist.

\textbf{Where do people go when they die?} To the barrows. Theona's dead are buried in long, low mounds that line the eastern cliffs. The mounds are planted with gorse and broom. The gorse blooms yellow in spring. The yellow is for mourning. The broom is for memory. The dead are not mourned. The dead are remembered. Remembering is enough.

\textbf{How do you apologize for a serious wrong?} You walk the Ninth Mile—a stretch of road that circles the central isle. The road is nine miles long. You must walk it at night, alone, with no light. If you reach the end by dawn, you are forgiven. If you do not, the Host will find you. The Host is always looking. The Host is always waiting.

\textbf{What is the most important festival?} The Feast of the Eighth Cup (autumn equinox). Every household pours eight cups of mead—one for each of the living generations. The ninth cup is poured onto the ground. The ground drinks. The ancestors listen. The ancestors do not answer.

\textbf{What is the secret that could destroy a powerful person?} The Three-Isles King is a changeling. The real king died in 798 AR, drowned in a peat bog while hunting. The changeling was placed by the Lady Beneath as part of a bargain. The bargain was for peace between the isles. The peace has held. The changeling has aged. The changeling's reflection does not.
\end{almanac}

\newpage
\input{../generators/theona.tex}
\newpage

% ------------------------------
% CHAPTER: ZAKOV
% ------------------------------
\chapter{Zakov}
\emph{Salt \& Serpent.}

\noindent
{\Large
\textit{The tide here keeps better ledgers than any scribe, and the reefs remember every secret whispered above their depths. Power flows like the currents—those who understand them rule the waves.}

\vspace{0.5em}

I am a Dock-Rat. My tattoo is of the Sunken Quarter at low tide. The buildings are upside down.

\vspace{0.5em}

The pirate-city of the Dolmis coast isn't just a den of thieves—it's a corporation with a flag, and oh, how the Salt Prince loves playing king while the Syndicate Silent pulls the strings. Here, every manifest is a lie, every harbor a doorstep to a dozen conspiracies, and every fog-light a lure or a warning. The city breathes with the sea: at low tide, secrets exposed; at flood, the drowned walk the Sunken Quarter and knock on doors that should be locked.

\vspace{0.5em}

The Salt Prince's crown floats. The Syndicate's ledger is written in disappearing ink. The Kraken's Tongue lives in a bell tower that only exists at high tide. None of this is metaphor. All of it is true. All of it wants something from you.

\vspace{0.5em}

Never take a coin that tastes of brine. The Serpent's Mark is not a blessing. It is a debt that drowns.

\vspace{1em}

\noindent\textit{— Ravel, Dock-Rat, who has never slept above the high-water mark}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salted fish and hard cheese—the fish is dried on the docks, the cheese is smuggled from the mainland. The cheese is always stale. The fish is always fresh. The contradiction is never explained.

\textbf{What does a child learn before they can walk?} To read the tide tables. Zakov children are taught the schedule of the sea before they learn their letters. A child who cannot read the tide is a child who will be caught by the flood. The flood is not water. The flood is the Syndicate. The Syndicate collects debts at high tide.

\textbf{What is the first thing you notice about a stranger?} Their coin-pouch. Zakov runs on silver—good silver, bad silver, silver that has been under the sea. A stranger whose pouch clinks with fresh-minted coin is a merchant. A stranger whose pouch is full of salt-crusted bits is a pirate. A stranger whose pouch is empty is a liar. Liars do not last long in Zakov. The sea takes them.

\textbf{What is the most common crime?} Piracy—but piracy is not a crime, it is a profession. The Syndicate licenses privateers. The privateers pay a tithe. The tithe funds the harbor. The harbor protects the ships. The ships are protected from pirates. The pirates are the Syndicate. The cycle is beautiful.

\textbf{What does no one talk about but everyone knows?} The Salt Prince is not a person. The title is held by a ring—a band of black coral that has been passed from captain to captain for three centuries. The ring is cursed. The ring's wearer cannot sleep. The ring's wearer cannot die. The ring's wearer cannot leave Zakov. The current Prince has worn the ring for forty years. He has not slept in forty years. He has not smiled in forty years. He has not left the palace in forty years.

\textbf{Where do people go when they die?} To the Sunken Quarter—the drowned district that lies beneath the harbor. The dead are weighted with chains and dropped into the canals. The canals empty into the sea. The sea returns the bodies at low tide. The bodies are collected by the Syndicate. The Syndicate sells them to the Salt-Prince's collectors. The collectors use them for—no one asks. No one wants to know.

\textbf{How do you apologize for a serious wrong?} You pay the Salt Tithe—a silver coin for every year of the wronged party's life. The coin must be a salt coin—one that has been under the sea for a full tide cycle. The coin is thrown into the harbor. If it sinks, you are forgiven. If it floats, you are not. If it floats back to shore, you are cursed. The curse is the Serpent's Mark. The Mark is a debt that drowns.

\textbf{What is the most important festival?} The Rising of the Tides (summer solstice). The harbor fills with luminescent algae. The algae glows green. The green is called "the Kraken's Eye." The Kraken is said to rise on this night. The Kraken has not risen in living memory. The Kraken is patient. The Kraken is waiting.

\textbf{What is the secret that could destroy a powerful person?} The Pirate Queen is a spy. She was planted by the Kahfagian Admiralty forty years ago. Her mission was to destabilize the Syndicate. She has become the Syndicate. The Admiralty does not know. The Admiralty still sends her orders. She burns them. The ashes are mixed with her tea.
\end{almanac}

\newpage
\input{../generators/zakov.tex}
\newpage

% ------------------------------
% CHAPTER: AELINNEL (GNOMES)
% ------------------------------
\chapter{Aelinnel (Aevrossa)}
\emph{Stone, Bough, and Bright Things.}

\noindent
{\Large
\textit{The law here flows like water over stone—it finds its own path, and woe to those who try to dam it with paper and pretense. The earth itself will testify against the unjust.}

\vspace{0.5em}

I am a Charcoal-Burner. My kilns tell us more than your books ever could. When the smoke curls left, the tide's with you. When it splits in three, best stay by your own hearth till it burns straight again.

\vspace{0.5em}

The gnomish homeland isn't just a place where they count steps—it's a place where they worship the count, and oh, how the Aelinnel love a good loophole in their beloved numerology. They count everything: steps, bells, offerings—because to them, precision isn't just virtue; it's leverage with a supernatural clause.

\vspace{0.5em}

The fae courts of the Thorn Lady and the Green Knight watch every threshold. The Stone Prince keeps the old pacts indexed in ledgers that breathe. And the Green Gate opens at neap tide, expecting exact change in truths. The Aelinnel don't just live between worlds—they arbitrate them.

\vspace{0.5em}

Never bring naked iron past a hawthorn arch unless it is gilded or named. Copper is polite, silver is opinionated, cold iron is an accusation. The hawthorn will remember which you carried.

\vspace{1em}

\noindent\textit{— Tallow, Charcoal-Burner, who has never lost a sheep to the Hollow but has lost three helpers who did not listen}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Honey-cakes and mushroom tea—the honey is from the hawthorn groves, the mushrooms are harvested from the root-caverns. The cakes are baked in copper moulds. The moulds are shaped like animals. The animals are not edible. The shapes are for luck.

\textbf{What does a child learn before they can walk?} To count in eights. Aelinnel children are taught to group everything in eights—steps, breaths, heartbeats. The ninth is not a number. The ninth is a gap. The gap is where the Hollow sits. The Hollow is not a monster. The Hollow is a creditor. The debt is never paid in full.

\textbf{What is the first thing you notice about a stranger?} Their tools. Aelinnel carry copper—spoons, chisels, measuring rods. A stranger whose tools are iron is a barbarian. A stranger whose tools are silver is a priest. A stranger whose tools are gold is a fool. Gold does not ring true. Gold is for bribes. Bribes are for secrets.

\textbf{What is the most common crime?} Miscounting—but miscounting is not a crime, it is a geometric error. The error is corrected by the Stone Prince's ledgers. The ledgers are kept in the Hall of Sums. The Hall is guarded by clockwork sentinels. The sentinels do not sleep. The sentinels do not forgive. The sentinels count every breath.

\textbf{What does no one talk about but everyone knows?} The Stone Prince is not a prince. The title is held by a committee of six elders who have never been seen together. They communicate through copper messages left in the Hall of Sums. The messages are written in a code that changes with the tides. The code has not been broken in three centuries. The code is not meant to be broken. The code is meant to be trusted.

\textbf{Where do people go when they die?} To the tide-gates—stone arches at the edge of the sea. The dead are laid on the threshold. The tide comes in. The tide takes them. The gates are older than the Aelinnel. The gates were built by something else. The gates are still watching.

\textbf{How do you apologize for a serious wrong?} You offer a counting—a public recitation of your wrong in the Stone Prince's court. The court is a circle of copper chairs. The chairs are cold. The wrongdoer stands in the centre. The elders count the syllables of your apology. If the count is even, you are forgiven. If the count is odd, you must leave an offering. The offering is measured in copper. The copper is weighed against your breath. The scales are never balanced.

\textbf{What is the most important festival?} The Opening of the Green Gate (autumn equinox). The gate appears at low tide, a stone arch that leads to—no one knows. The Aelinnel do not enter. They leave offerings: copper, honey, a single true name. The gate accepts. The gate closes. The gate waits.

\textbf{What is the secret that could destroy a powerful person?} The Green Knight is not a knight. She is the gatekeeper—the last survivor of the people who built the tide-gates. She wears a helm of polished brass. Her face has not been seen in living memory. Her voice is a recording. Her sword is a key. The key opens the gate. The gate does not want to be opened.
\end{almanac}

\newpage
\input{../generators/aelinnel.tex}
\newpage

% ------------------------------
% CHAPTER: AELAEREM (HALFLINGS)
% ------------------------------
\chapter{Aelaerem}
\emph{Hearth \& Hollow.}

\noindent
{\Large
\textit{A proper harvest feast requires three things: the finest cider aged in the right barrel, a guest-loaf baked with intention, and the wisdom to know when to invite the quiet ones to sit by the fire.}

\vspace{0.5em}

I am a Miller. My geese know a stranger's step from a mile off, and they've never been wrong. One is named "Stranger." It has never been wrong.

\vspace{0.5em}

The halfling downs aren't just a place where they keep the Hollow at bay—they're a marketplace for fear, and oh, how the Aelaerem love a good bargain on safety. They use red threads and offerings not out of piety, but because they know the Hollow's attention is a commodity—and they'll pay dearly to avoid it.

\vspace{0.5em}

A red door means bread, salt, and one safe night if you come honest. A cup-mark stone on the verge accepts butter as a prayer. The watch-geese know your step before you do, and the scarecrow turns to face the road when the Hollow is hungry. The Neighbors—the things under the barrows—are not enemies. They are neighbours. And neighbours, if treated politely, will stay on their side of the hedge.

\vspace{0.5em}

Never refuse an offered cup. The Hollow counts refusals as invitations. Pour the ninth cup for the empty chair, and you may feed it—or feed it.

\vspace{1em}

\noindent\textit{— Bram, Miller of the Watch-Geese, who has never been surprised}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Porridge with honey and a pinch of salt—the salt is for the Hollow. Every Aelaerem household leaves a pinch on the threshold each morning. The Hollow does not eat. The Hollow counts. The salt is an offering. The offering is a reminder.

\textbf{What does a child learn before they can walk?} To tie a red thread. Aelaerem children are given a length of red wool at birth. By age three, they can tie the eight hearth-knots: welcome, farewell, protection, memory, warning, promise, apology, and the ninth—the one that is never tied. The ninth knot is for the Hollow. The Hollow does not tie knots. The Hollow waits for you to tie them yourself.

\textbf{What is the first thing you notice about a stranger?} Their shadow. Aelaerem shadows are long and thin, even at noon. A stranger whose shadow is short has been touched by the Neighbors. A stranger whose shadow is wrong—too many limbs, too few—is a Neighbor. Do not invite them in. Do not refuse them bread. The law is clear: feed them, but do not ask their name.

\textbf{What is the most common crime?} Offering the ninth cup—but the ninth cup is not a crime, it is a negotiation. The pourer offers a drink to the empty chair. The Hollow may accept. If it does, the house is blessed for a year. If it does not, the house is marked. The mark is a red thread tied to the door. The thread is never removed. The house is never sold.

\textbf{What does no one talk about but everyone knows?} The Apple-Matron is a title, not a person. The Matron is the oldest woman in the village. When she dies, the second-oldest takes her place. The Matron's memory is the village's memory. The Matron forgets nothing. The Matron's memory is fading. The village is forgetting. The forgetting is the Hollow's doing.

\textbf{Where do people go when they die?} To the barrows—the low, green mounds that dot the downs. The dead are buried in family plots, marked with cup-stones. The cup-stones are filled with rainwater. The rainwater is changed weekly. The water is for the dead. The dead are not thirsty. The dead are counting.

\textbf{How do you apologize for a serious wrong?} You leave a red thread on the wronged party's doorstep. The thread must be tied in a specific knot—the knot of apology, which is also the knot of debt. The wronged party may accept the thread or cut it. If they accept, the debt is recorded. If they cut it, the debt is unpaid. The unpaid debt grows. The growing debt attracts the Hollow.

\textbf{What is the most important festival?} The Harvest Home (autumn). The last sheaf of grain is tied with red thread and left in the field. The Hollow takes it. The Hollow does not take anything else. The Hollow is patient. The Hollow will wait for the ninth cup.

\textbf{What is the secret that could destroy a powerful person?} The Pale Shepherd is not a shepherd. It is a title held by the Hollow itself. The Shepherd appears as a figure in grey robes. Its face is a mask of bone. Its voice is a recording of the last words of the person who saw it last. That person is always dead. The recording is always accurate.
\end{almanac}

\newpage
\input{../generators/aelaerem.tex}
\newpage

% ------------------------------
% CHAPTER: THE VALEWOOD
% ------------------------------
\chapter{The Valewood}
\emph{Empire Under Leaves.}

\noindent
{\Large
\textit{The forest here remembers empires that crumbled to dust, and its laws are older than your kingdoms. To walk these paths is to enter into covenant with powers that measure time in seasons, not years.}

\vspace{0.5em}

I am a Pathweaver. My cord has eight knots. The ninth is the one she never ties.

\vspace{0.5em}

The Valewood isn't just ancient—it's a graveyard of empires, and oh, how the Lethai-ar ghosts love a good deal on forgotten magic. They don't just rule as echoes; they trade in them. The air hums with half-remembered songs, and the ruins bleed their mysteries into root and stream. Fae courts linger in clearings that shift with moonlight, beast-kin patrol moss-choked cairns, and echoes of imperial shades still march patrols on vanished streets.

\vspace{0.5em}

The Hazel Queen speaks only in conditional tense. The Alder King's crown is a noose of thorns. The Huntsman's face is a stag's skull, and his voice is the wind in the leaves. The forest does not forgive arrogance. It expects humility, patience, and the willingness to trade a piece of yourself for safe passage.

\vspace{0.5em}

Never eat the fruit of a tree that has a face. The forest does not forget its debtors, and the first bite is always the price of admission. The second bite is the interest.

\vspace{1em}

\noindent\textit{— Pathweaver Sila, who knows every star-road but will not speak their names}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Acorn bread and wild honey—the acorns are leached in stream-water for three days, the honey is taken from hives in the high branches. The bees are sacred. The bees are not to be harmed. A beekeeper who harms a bee is cursed. The curse is the stinging. The stinging is endless.

\textbf{What does a child learn before they can walk?} To name the trees. Valewood children are taught the names of the eight sacred groves: Hazel, Alder, Oak, Rowan, Ash, Thorn, Yew, and the ninth—the one that has no name. The ninth grove is where the Hollow sleeps. The Hollow is not to be woken.

\textbf{What is the first thing you notice about a stranger?} Their way-cord. Every Valewood traveler carries a knotted cord—one knot for each mile they have walked. A stranger whose cord has nine knots has walked the forbidden path. A stranger whose cord is knotted with red thread is a Pathweaver. A stranger with no cord has no way. The forest will find them. The forest will lead them in circles until they forget why they came.

\textbf{What is the most common crime?} Breaking the star-road—but the star-roads are not roads, they are agreements. A traveler who strays from the path is a traveler who breaks the covenant. The covenant is older than the Valewood. The covenant was made between the first trees and the first stars. The stars do not forgive. The trees do not forget.

\textbf{What does no one talk about but everyone knows?} The Hazel Queen is not the queen. She is the voice of the forest—a title that passes from dryad to dryad with each turning of the season. The current Queen has held the title for three centuries. She is not a dryad. She is a ghost—the last survivor of the empire that fell beneath the leaves. She remembers the cities. She remembers the stones. She remembers the names of the dead. The dead do not remember her.

\textbf{Where do people go when they die?} To the star-roads—the paths that lead between the trees and between the worlds. The dead are not buried. They are laid on stone altars at the edges of the groves. The stars pull them upward. The trees pull them downward. The dead choose. The dead always choose the trees. The trees remember.

\textbf{How do you apologize for a serious wrong?} You offer a name—your own or another's. The forest accepts the name. The forest weaves it into the bark of the oldest oak. The oak is called the Witness. The Witness has stood for nine thousand years. The Witness has never forgotten a single name. The Witness will not forget yours.

\textbf{What is the most important festival?} The Turning of the Leaves (autumn). The trees shed their colours in a single day. The leaves carpet the forest floor. The leaves are gathered and burned in the high clearings. The smoke carries the year's debts to the stars. The stars answer with light. The light is forgiveness. The light is also a warning.

\textbf{What is the secret that could destroy a powerful person?} The Alder King is not a king. He is the forest's warden—a title passed from fallen empire to fallen empire. The current King is a Lethai-ar who was born in the year the last empire fell. He has not aged since. He does not eat. He does not drink. He does not sleep. He walks the star-roads at night. The star-roads are empty. The dead have stopped walking.
\end{almanac}

\newpage
\input{../generators/valewood.tex}
\newpage

% ============================================================
% PART IV: THE SOUTHERN REALMS (AKILAN)
% ============================================================
\part{The Southern Realms (Akilan)}
% After \part{The Southern Realms (Akilan)}
\begin{quote}
  \small
  \textit{``The ledger does not forgive arithmetic errors. A single miscounted bushel can unmake a house, a tribe, an empire. I am not here to be kind. I am here to be correct. But even the Widow's ink runs when the monsoon weeps.''}
  \noindent — Dusana of the Raven Road, at the weigh‑house of Mbaro
  \end{quote}

% ------------------------------
% CHAPTER: OSHIIRA
% ------------------------------
\chapter{Oshiira}
\emph{The Ledger Empire.}

\noindent
{\Large
\textit{The ledger does not forgive arithmetic errors. A single miscounted bushel can unmake a house, a tribe, an empire. I am not here to be kind. I am here to be correct.}

\vspace{0.5em}

I am a Grain Inspector. My stamp leaves a mark that glows under moonlight. I know.

\vspace{0.5em}

Oshiira is a confederacy of hundreds of tribes and city-states, united by the Grain Ledger. The capital, Mbaro, is a city of canals, weigh-houses, and the great Granary Tower. Oshiira is the breadbasket of the Amaranthine; its millet and sorghum fill the holds of Ecktorian, Vhasian, and Silkstrand ships. The empire operates through a bureaucracy as complex as any in the north—weighing-houses, canal-mothers, and ledgers that track every grain shipment with mathematical precision. But here's the secret the Widow doesn't want you to know: the Ledger has a page that has been torn out. That page named the true heir. No one knows where it is. Everyone is looking.

\vspace{0.5em}

The Vermilion Corps audits not only grain but also loyalty. The Narrow Captains provoke Kahfagia just to remind the Admiralty who really controls the straits. And the Widow—the empress who has held the confederacy together for centuries—is dying. Paranoia is slowly replacing wisdom. Her four children circle the throne like vultures. The Vermilion Corps maintains order, but no one knows which heir the Corps truly serves. Not even the Corps itself.

\vspace{0.5em}

Never speak of succession before naming the Widow. The empress's health is a state secret, and the Vermilion Corps has ears in every weigh-house. Speak carelessly, and your name may be the next one erased from the Ledger.

\vspace{1em}

\noindent\textit{— Vermilion Corps Auditor, who counts her own breaths to ensure she is not lying}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Millet porridge with dried fish—the millet is grown in the flood plains of the Sahel, the fish is smoked over canal-barges. The porridge is always the same. The fish varies by season. The seasons are measured by the height of the river. The river is the Widow's blood. The Widow's blood is low.

\textbf{What does a child learn before they can walk?} To weigh a grain. Oshiira children are given a balance scale and a single grain of millet. By age five, they can tell you the weight difference between a grain harvested in the northern Sahel and a grain harvested in the southern riverlands. The difference is the tax. The tax feeds the empire. The empire is hungry.

\textbf{What is the first thing you notice about a stranger?} Their hands. Oshiira merchants have ink-stained fingers—the ink is from the Ledger. A stranger whose fingers are clean has never signed a contract. A stranger whose fingers are stained black has signed too many. A stranger whose fingers are stained red has signed a death warrant. The red ink is not ink. The red ink is blood.

\textbf{What is the most common crime?} Ledger fraud—but ledger fraud is not a crime, it is a career. The Vermilion Corps audits the auditors. The auditors audit the clerks. The clerks audit the merchants. The merchants audit the farmers. The farmers audit the land. The land does not lie. The land does not need to.

\textbf{What does no one talk about but everyone knows?} The Widow has not been seen in public for seven years. Her voice is a recording. Her seal is a stamp. Her decrees are written by the Vermilion Corps. The Corps signs in her name. The Corps speaks in her voice. The Corps is the Widow. The Widow is dead. The empire continues.

\textbf{Where do people go when they die?} To the Granary Tower—the great stone silo at the centre of Mbaro. The dead are not buried. They are ground into meal and mixed with the millet. The grain is shipped across the Amaranthine. The dead feed the empire. The empire does not know. The empire does not ask.

\textbf{How do you apologize for a serious wrong?} You submit a correction to the Ledger. The correction must be witnessed by a Vermilion Corps auditor. The auditor stamps the correction with their seal. The seal glows under moonlight. The moonlight is the Widow's eye. The Widow sees everything. The Widow forgives nothing.

\textbf{What is the most important festival?} The Harvest of the Ledger (autumn). The Ledger is opened in the great square. The names of the year's debtors are read aloud. The debtors are given until dawn to pay. The debtors who cannot pay are added to the Ledger's ninth page. The ninth page is sealed with wax. The wax is black. The black is mourning.

\textbf{What is the secret that could destroy a powerful person?} The Ledger's missing page is not missing. It was burned by the Widow herself, in the year the last heir died. The page's ash was mixed into the mortar of the Granary Tower. The ash still glows under moonlight. The glow spells the heir's name. The name is the Widow's own. The Widow killed her child. The empire does not know. The empire would not care.
\end{almanac}

\newpage
\input{../generators/oshiira.tex}
\newpage

% ------------------------------
% CHAPTER: ASHAAN
% ------------------------------
\chapter{Ashaan}
\emph{The Veiled Throne.}

\noindent
{\Large
\textit{The Veil is not a choice. It is survival. You will show your eyes and nothing more. Speak in thirds—offer two options and a hidden third. The ninth word is always a knife.}

\vspace{0.5em}

I am a Breath-less Agent. They say Ikasha sleeps. I say she is patient. The Sisters have ruled for generations; they believe the Veil is eternal. But the Veil is silk. Silk frays. And I am learning to bite through—one whisper at a time.

\vspace{0.5em}

Ashaan is a nation that died long ago. The Veiled Sisters keep its corpse walking through spirit-binding, terror, and the endless consumption of their own people. The capital, Galanina, is a city of black marble and brass cages. The Red Desert, which now covers half the country, was once fertile land; the Sisters' magic leeched the life from it. There is no recovery, only a slower dying. But here's the secret the Sisters don't want you to know: the Veil is cracking. The spirits they've bound for centuries are whispering to each other. And the Hollow has begun to notice.

\vspace{0.5em}

The Esoiti control every ledger and every gate. The Black Hand hunts without uniform—but the Hand is human. They were once bakers, clerks, mothers, soldiers. Some are fanatics who chose the knife. Others are desperate souls who were given a choice: serve or vanish. They do not sleep well. Their hands shake when the Sisters’ acolytes pass. And the Sisters—the Sisters are not human. They do not negotiate. They do not threaten. They state what will be. Their voices are the same as the wind over a mass grave. You cannot bribe a divine disaster. You can only survive its attention.

The Breath-less meet in root cellars, learning to read ledgers without leaving marks. We are not heroes. We are the ones who learned that a whisper can be louder than a scream—and that the Sisters fear silence more than they fear rebellion. Silence means the Veil is healing. Silence means the bound are forgetting. The Sisters cannot abide forgetting.

\vspace{0.5em}

Never name the dead. The Sisters bind spirits to names. Speak a dead name aloud, and you may find yourself bound alongside them.

\vspace{1em}

\noindent\textit{— Tamar, Breath-less cell leader, who has three names—each for a different jurisdiction}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Black bread and thin soup—the bread is made from grain grown on the edge of the Red Desert, the soup is flavoured with salt and memory. The memory is of green fields. The fields are gone. The bread is bitter.

\textbf{What does a child learn before they can walk?} To veil their face. Ashaan children are taught to cover their mouths when speaking, their eyes when listening, their ears when hearing secrets. A child who cannot veil is a child who will be taken by the Sisters. The Sisters take children. The children become spirits. The spirits serve the Sisters. The Sisters never stop taking.

\textbf{What is the first thing you notice about a stranger?} Their veil. Ashaan veils are coded—red for mourning, white for marriage, black for the Sisters. A stranger whose veil is grey is a Breath-less agent. A stranger whose veil is torn is a fugitive. A stranger with no veil is a fool. Fools are collected by the Black Hand. The Black Hand does not ask questions. The Black Hand was once a person. The Black Hand no longer remembers who.

\textbf{What is the most common crime?} Spirit-trading—but spirit-trading is not a crime, it is the economy. The Sisters bind spirits to brass cages. The cages are sold to the Esoiti. The Esoiti sell the spirits to foreign buyers. The foreign buyers do not know what they are buying. The Sisters do not care. The Sisters are not merchants. They are a natural disaster with a ledger.

\textbf{What does no one talk about but everyone knows?} \textit{Everyone is afraid.} The Black Hand fears the Esoiti’s audits. The Esoiti fears the Sisters’ gaze. The Sisters fear the Hollow’s patience. The baker fears the guard. The guard fears the Hand. The Hand fears the name on tomorrow’s list. The only person in Ashaan who is not afraid is the child who has not yet learned to read the ledger. That child will be afraid by next spring.

\textbf{Where do people go when they die?} To the brass cages. The dead are not buried. They are bound into spirit-lanterns and hung from the walls of Galanina. The lanterns glow green. The green is the light of the bound. The bound cannot speak. The bound cannot scream. The bound can only watch. Sometimes they weep. The tears evaporate into glass.

\textbf{How do you apologize for a serious wrong?} You offer a name—a dead name, a forgotten name, a name the Sisters do not yet know. The Sisters bind the name to a brass cage. The cage is placed in the Hall of Silence. The Hall is full. The Hall is always full. The Hall is never silent. The cages whisper. The whispers are the names of those who were erased.

\textbf{What is the most important festival?} The Night of the Unbound (winter solstice). The spirits strain against their cages. The Sisters renew their bindings. The ceremony takes three days. The city is silent for three days. The silence is not prayer. The silence is fear. If you listen closely during the silence, you can hear the Hollow breathing.

\textbf{What is the secret that could destroy a powerful person?} The High Sister is not a sister. She is a spirit—the first one bound, nine hundred years ago. She bound herself to the throne. She cannot leave. She cannot die. She cannot stop binding. The other Sisters do not know. The other Sisters are her prisoners. The prisoners guard the jailer. The jailer does not sleep. Her veil is made of gold coins from a dozen dead empires. The coins do not clink. They weep when she speaks.
\end{almanac}

\newpage
\input{../generators/ashaan.tex}
\newpage

% ------------------------------
% CHAPTER: NGOMEBE
% ------------------------------
\chapter{Ngomebe}
\emph{The Stone Men.}

\noindent
{\Large
\textit{A city that stops moving is a city that has chosen to die. The stone remembers every mile, every roller, every oath sworn in the shadow of the keystone. We do not build cities. We raise them from the earth and teach them to walk.}

\vspace{0.5em}

I am a Stone-Cutter. My grandfather carved the keystone for this city. My father greased the rollers. I inherited a sledge and a song. The song is the city's heartbeat. If I stop singing, the city stops walking—and so do I.

\vspace{0.5em}

Ngomebe are a people of deep, iron-dark skin and older ties to the southern ranges. Their cities walk—colossal stone structures on rollers and sledges, dragged by teams of oxen, magic, and Aeler-designed capstans. A walking city is not a curiosity; it is a living argument against stillness. But here's the secret the Mason-Kin don't want you to know: the stone is not just a material. It listens. It remembers. And sometimes, it disagrees. When the keystone cracks, it is not a flaw—it is a message. The ancestors are trying to tell you something. The question is whether you're brave enough to read the cracks.

\vspace{0.5em}

The Root-binders plant acorns in the roller-pits, singing to the earth. They believe the cities should stop and let the forest reclaim the stone. The Aeler engineers maintain the bearings, but they are neither masters nor servants—they are partners in a pact older than memory. And the Ancestral Healers speak for the dead, who vote through the cracks in the keystone. Their votes are always unanimous. Their verdicts are always final.

\vspace{0.5em}

Never carve a name into living rock. The stone remembers. It will whisper that name to every spirit of the earth, and the debt will follow your bloodline for seven generations.

\vspace{1em}

\noindent\textit{— Kaelen, Stone-Cutter, whose city left him behind}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Stone-ground meal and fermented milk—the meal is ground between the city's own millstones, the milk is from goats that walk alongside the city. The goats are older than the city. The goats remember where the city has been. The goats do not speak. The goats do not need to.

\textbf{What does a child learn before they can walk?} To sing the city's song. Ngomebe children are taught the rhythm of the rollers—eight beats to a mile, eight miles to a day, eight days to a season. The ninth beat is the city's secret. The ninth beat is never sung aloud.

\textbf{What is the first thing you notice about a stranger?} Their dust. Ngomebe are coated in fine stone-dust—from the quarries, the rollers, the wind that follows the city. A stranger whose dust is white has come from the mountains. A stranger whose dust is red has come from the desert. A stranger with no dust has never walked with a city. The cities do not trust the dustless.

\textbf{What is the most common crime?} Stone-theft—but stone-theft is not a crime, it is a wound. The stone is the city's memory. A stolen stone is a forgotten year. The year is forgotten by everyone who walked it. The forgetfulness spreads. The city's song stutters. The stutter is a warning.

\textbf{What does no one talk about but everyone knows?} The city is not moving toward anything. It is moving away. The Ngomebe are fleeing something that lies to the south. The thing in the south is older than the stone. The thing in the south does not walk. It waits.

\textbf{Where do people go when they die?} Into the keystone. The dead are not buried. Their ashes are mixed into mortar and pressed into the city's foundation. The dead become part of the city's weight. The city walks heavier each year. The dead do not complain. The dead are the city's anchor.

\textbf{How do you apologize for a serious wrong?} You carve a new roller. The roller must be cut from a single stone, with no cracks, no flaws, no forgotten names. The roller is placed beneath the city's foundation. If it rolls true, you are forgiven. If it cracks, you are not. If it crumbles, you are added to the mortar.

\textbf{What is the most important festival?} The Turning of the Rollers (spring equinox). The city stops for one day. The rollers are inspected, greased, and blessed. The city's song is sung from dawn to dusk. The song has eight verses. The ninth verse is silence. The silence is for the dead. The dead are listening.

\textbf{What is the secret that could destroy a powerful person?} The city is not a city. It is a tomb—a monument to a king who died before the first stone was laid. The king's name is carved into the keystone. The keystone is at the city's heart. The heart is inaccessible. The king is patient. The king is waiting to be remembered.
\end{almanac}

\newpage
\input{../generators/ngomebe.tex}
\newpage

% ------------------------------
% CHAPTER: TAHARKA
% ------------------------------
\chapter{Taharka}
\emph{The Monsoon Crown.}

\noindent
{\Large
\textit{The passes remember every army that tried to climb them. The mountains do not forgive haste, and they do not forget a foreign banner. We do not guard the highlands. We listen to them. They tell us who to let through.}

\vspace{0.5em}

I am a Horse Trader. Before you praise the rider, praise the horse. Before you praise the horse, praise the mountain that grew its lungs. A Taharkan foal is born breathing thin air. That is why our cavalry has never been broken.

\vspace{0.5em}

Taharka grows rich on monsoon trade—spices, coffee, and incense that travel east to the distant ocean and west to the Gulf. The kingdom has never been conquered; its passes are a graveyard for Ashaani armies and a barrier that Oshiiran grain barges cannot cross. The highlands are a fortress of rock and thin air, where every valley speaks a different dialect and every ridge has a watchtower. But here's the secret the Crown Council doesn't want you to know: the passes are alive. They have a memory. They remember every army that tried to climb them, every foreign banner that flew too high, every oath that was broken at the summit. And they do not forgive.

\vspace{0.5em}

The Ridge-Watchers are the eyes of the kingdom, stationed in rock-chapels that double as signal towers. The horse is not a mount; it is a lineage, a treaty, a weapon. To control the highland breeds is to control the passes—and to control the passes is to remain free. The lowland coast operates under a different law: the ujamaa of the dhows, a mutual aid network where every captain owes a share of every voyage to the community. A captain who cheats the ujamaa finds his ship's name struck from every mooring post in the archipelago.

\vspace{0.5em}

Never refuse a second cup of tea. The first is hospitality; the second is a test of honour. The third is a promise. Refuse the second, and you have insulted the house—and the house's ancestors, who are listening through the floorboards.

\vspace{1em}

\noindent\textit{— Ridge-Watcher Captain Anwar, who has not descended below the snow line in twenty years}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Barley flatbread and spiced tea—the bread is baked on iron griddles, the tea is brewed with cardamom and mountain herbs. The herbs are harvested from passes that only the Ridge-Watchers know. The Ridge-Watchers do not share.

\textbf{What does a child learn before they can walk?} To read the wind. Taharkan children are taught the eight winds by name: North, South, East, West, Monsoon, Dry, Storm, and the ninth—the wind that has no name. The ninth wind is the mountain's breath. The mountain's breath is death.

\textbf{What is the first thing you notice about a stranger?} Their breath. The air is thin in Taharka. A stranger who breathes too fast has not acclimated. A stranger who breathes too slow is hiding something. A stranger who does not breathe at all is a spirit. The spirits are not to be spoken to.

\textbf{What is the most common crime?} Horse theft—but horse theft is not a crime, it is a declaration of war. The horse is the kingdom's bloodline. A stolen horse is a stolen ancestor. The thief's clan is responsible. The clan pays in blood or in trade. The trade is never enough. The blood is always too much.

\textbf{What does no one talk about but everyone knows?} The Crown Council has not met in ten years. The passes have been closed by snow, by war, by the simple fact that the councillors cannot agree on a meeting place. The kingdom is ruled by the Ridge-Watchers. The Ridge-Watchers are not elected. The Ridge-Watchers are the mountains' voice. The mountains are silent.

\textbf{Where do people go when they die?} To the high passes. The dead are laid on stone platforms, wrapped in woollen shrouds. The wind takes them. The wind carries their names to the peaks. The peaks remember. The peaks are the kingdom's cemetery.

\textbf{How do you apologize for a serious wrong?} You offer a horse—the best horse in your stable. The horse is given to the wronged party. If they accept, the debt is paid. If they refuse, the debt is doubled. If they kill the horse, the debt is blood. Blood is never repaid.

\textbf{What is the most important festival?} The First Rain (monsoon's arrival). The passes open for the first time in months. The caravans descend from the highlands. The lowland ports receive the first spice ships. The celebration lasts nine days. The ninth day is silent. The silence is for the dead of the passes.

\textbf{What is the secret that could destroy a powerful person?} The Crown Council's missing member is not missing. He is buried under a landslide that was not an accident. The landslide was caused by the Ridge-Watchers. The councillor had learned the passes' secret: they are not natural. They were carved by something older. The something is still there. The Ridge-Watchers are its servants.
\end{almanac}

\newpage
\input{../generators/taharka.tex}
\newpage

% ------------------------------
% CHAPTER: SEKOGO
% ------------------------------
\chapter{Sekogo}
\emph{The Green Council.}

\noindent
{\Large
\textit{The jungle has an immune system. When a fever rises, we cut it out. When an oath is broken, we collect the price. We are not cruel. We are necessary. The tree remembers every unpaid debt, and we are its teeth.}

\vspace{0.5em}

I am a Honey-Gatherer. The bees know my name. They do not sting me because I sing to them the song of the first blossom. My grandmother taught me that the honey is not a gift; it is a loan. One day, the grove will ask for it back.

\vspace{0.5em}

Sekogo lies between the Sahel grasslands and the southern plateau—a dry savannah that grows wetter as you travel east, until the baobabs give way to deep jungle and the air turns thick with the smell of damp earth and overripe fruit. Here, the Green Council rules through consensus, honey-orchards are sacred, and the old spirits still walk. Slavery is forbidden; the jungle itself would rise against any slaver who crossed its border. But here's the secret the Council doesn't want you to know: the jungle has a favourite. It is not the Conservators who protect the old pacts. It is not the Wakers who want to rouse the jungle with fever-mists and bloodier rites. It is the Ukwe—the Speaking Tree—that decides whose debts are collected and whose are forgiven. And the Ukwe does not answer to the Council.

\vspace{0.5em}

The Leopard Society hunts those who break the deepest taboos. The Earth Society keeps the Calendar of Debt, a sacred list written on bark and buried at the Ukwe's roots. The nganga work with minkisi—power objects inhabited by spirits—hammering nails into carved figures to release magic. Each nail is a new debt to the spirit within. Trickster spirits—children of Ananse, the spider-trickster—weave illusions and swap debts for laughter. And in the oldest groves, the Ninki Nanka—a serpent with a mirror-back—slithers between roots, its reflection showing the lies you have told.

\vspace{0.5em}

Never cut a marked tree. Each such tree is a living witness. The bark grows over the names of debtors, but the wood remembers. Cut it, and the Ukwe will add your name to its ledger.

\vspace{1em}

\noindent\textit{— Hunter Okoro, Leopard Society mask-wearer, who never removes his mask}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Mashed yams and groundnut stew—the yams are grown in clearings that shift with the seasons, the groundnuts are harvested by monkey-tribes who speak the old tongue. The monkeys are not paid. The monkeys are remembered. Memory is currency.

\textbf{What does a child learn before they can walk?} To name the spirits. Sekogo children are taught the eight spirits of the grove: River, Tree, Stone, Rain, Fire, Wind, Honey, and the ninth—the spirit that has no name. The ninth spirit is the Ukwe's shadow. The shadow watches. The shadow never forgives.

\textbf{What is the first thing you notice about a stranger?} Their mask. Every Sekogo wears a mask—carved wood, painted with the symbols of their lineage. A stranger whose mask is cracked has broken a taboo. A stranger whose mask is blank has no lineage. A stranger with no mask is not a person. The Leopard Society hunts the maskless.

\textbf{What is the most common crime?} Debt-forgetting—but debt-forgetting is not a crime, it is a sickness. The Ukwe keeps the Calendar of Debt. A name that is forgotten is a debt that is unpaid. The unpaid debt grows roots. The roots strangle the debtor's house. The house collapses. The jungle reclaims it.

\textbf{What does no one talk about but everyone knows?} The Green Council does not rule. The Ukwe rules. The Council is the Ukwe's voice—eight speakers chosen by the tree's roots. The ninth seat is empty. The ninth speaker was the tree itself. The tree does not speak. The tree waits.

\textbf{Where do people go when they die?} To the Ukwe's roots. The dead are not buried. Their bodies are laid at the base of the Speaking Tree. The roots absorb them. The tree grows a new leaf for each soul. The leaves whisper the names of the dead. The whisper is the wind.

\textbf{How do you apologize for a serious wrong?} You hammer a nail into a nkisi—a power figure carved by the nganga. The nail represents your wrong. The spirit inside the nkisi accepts your apology. The nail rusts. When the nail is gone, your debt is paid. The nail never rusts.

\textbf{What is the most important festival?} The Honey Harvest (autumn). The bees are sung to sleep. The hives are opened. The honey is collected in copper bowls. The first bowl is offered to the Ukwe. The tree drinks. The tree does not thank.

\textbf{What is the secret that could destroy a powerful person?} The Ukwe is not a tree. It is a prison—the binding of a spirit that tried to consume the jungle nine centuries ago. The spirit is still alive. The tree is its cage. The roots are its chains. The Green Council is its jailers. The jailers have forgotten what they guard.
\end{almanac}

\newpage
\input{../generators/sekogo.tex}
\newpage

% ------------------------------
% CHAPTER: LETHAI-AR OF THE CRIMSON VALLEY
% ------------------------------
\chapter{Lethai-ar of the Crimson Valley}
\emph{Silk Bridges and Web-Law.}

\noindent
{\Large
\textit{The web-law is not written on parchment. It is woven into every bridge, every strand, every knot. When a promise is made, the valley remembers. When a promise is broken, the silk tightens. We are not judges. We are the loom.}

\vspace{0.5em}

I am a Mark-Cutter. Each line on my skin is a promise I witnessed. Each line is a debt I carry. When I die, they will cut these marks from my body and bury them at the Ukwe's roots. Then the tree will remember what I owed. I hope it is enough.

\vspace{0.5em}

The Crimson Valley is a deep jungle rift between Sekogo and Oshiira—a wound of red stone, carved by a river that runs rust-coloured after every rain. Here, the Lethai-ar guard their borders with fierce neutrality, their chestnut-brown skin blending with the canyon walls. Silk bridges span the gorges, woven from spider-sap and oath-hair, strong enough to hold a caravan and light enough to sway in the monsoon wind. But here's the secret the wardens don't want you to know: the web-law is not unwritten because it is ancient. It is unwritten because it changes. Every warden carries a different clause in memory. Together, they form a complete web—but the web is never complete for long. The valley is learning. And it is not learning to be kind.

\vspace{0.5em}

The Lethai-ar are adopted children of the old powers—not fallen, but chosen to carry memory. Each can name the specific mountain, river, or grove where their ancestor left the spirit world to become mortal. Their skin-tattoos are not decoration; they are a public ledger, written on skin, read by anyone who knows the phase of the crimson moon. A Lethai-ar who bears a Mark they cannot fulfil loses standing. A warden who cuts a bridge without cause is exiled.

\vspace{0.5em}

Never step onto a silk bridge without first naming a witness. The bridge will remember your footsteps. If you fall, it will ask why no one was watching.

\vspace{1em}

\noindent\textit{— Silk-Warden Anara, whose left arm is covered in Marks from wrist to shoulder}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge and fermented tea—the rice is grown on terraces cut into the canyon walls, the tea is aged in clay pots buried beneath the bridges. The tea is bitter. The bitterness is for memory.

\textbf{What does a child learn before they can walk?} To read a Mark. Lethai-ar children are taught the eight Mark-shapes: Promise, Debt, Witness, Apology, Interest, Cut, Forgive, and the ninth—the empty space where a Mark was erased. The erased Mark is a broken promise. The broken promise is a cut. The cut is a scar.

\textbf{What is the first thing you notice about a stranger?} Their Marks. A stranger whose Marks are fresh has made new promises. A stranger whose Marks are faded has unpaid debts. A stranger whose Marks are cut has been exiled. The exiled are not spoken to. The exiled are not remembered.

\textbf{What is the most common crime?} Bridge-cutting—but bridge-cutting is not a crime, it is a war. The bridge is a promise between two shores. Cutting it is breaking that promise. The valley responds. The valley tightens its web. The web is the law. The law is the valley.

\textbf{What does no one talk about but everyone knows?} The valley is not a place. It is a creature—a sleeping serpent whose spine is the canyon walls, whose scales are the silk bridges, whose heart is the rust-coloured river. The Lethai-ar are its parasites. They live on its back. They weave its silk. They do not know it is asleep.

\textbf{Where do people go when they die?} To the river. The dead are not buried. Their bodies are wrapped in silk and cast into the rust-coloured water. The river carries them downstream, into the Ukwe's territory. The Ukwe drinks. The Ukwe remembers. The Ukwe does not forget.

\textbf{How do you apologize for a serious wrong?} You offer a Mark—a promise cut from your own skin. The Mark is pressed into the wronged party's ledger. If they accept, the debt is transferred. If they refuse, the Mark is buried. Buried Marks are debts that never die.

\textbf{What is the most important festival?} The Weaving of the New Bridge (spring). The silk is harvested from the canyon's spiders. The thread is spun by the wardens. The bridge is woven in a single night. The bridge is named after the year. The year is named after a promise. The promise is always the same: the valley will hold.

\textbf{What is the secret that could destroy a powerful person?} The river's rust colour is not iron. It is blood—the blood of the serpent that sleeps beneath the canyon. The serpent is wounded. The wound has been bleeding for nine centuries. The Lethai-ar know. The Lethai-ar do nothing. The serpent's dream is the valley's law. If the serpent wakes, the law ends.
\end{almanac}

\newpage
\input{../generators/crimson-valley.tex}
\newpage

% ============================================================
% PART V: THE WAY OF SILK
% ============================================================
\part{The Way of Silk}

% After \part{The Way of Silk}
\begin{quote}
  \small
  \textit{``The road is a ledger. Every oasis is a credit, every broken wheel a debt. The Tulkani tie knots to remember who owes whom. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. And the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.''}
  \noindent — Dusana of the Raven Road, at a caravanserai whose name she refused to give
  \end{quote}
% ------------------------------
% CHAPTER: FHARA
% ------------------------------
\chapter{Fhara}
\emph{The Oasis Courts.}

\noindent
{\Large
\textit{Water is not a right. It is a loan from the sand. The oasis remembers every drop drawn, every cup poured, every promise made at its edge. I do not judge the living. I judge the debt.}

\vspace{0.5em}

I am a Water-Clerk. My grandfather dug this well with a shovel and a prayer. My mother sealed the walls with clay from the dry river. I sit in the shade and measure every bucket. The sand does not forgive a miscount.

\vspace{0.5em}

Water isn't just law in Fhara—it's a market, and oh, how the water-clerks love a good deal on aquifer access. They don't just negotiate in three cups; they auction the right to the fourth cup. The Fhara are merchants, poets, and negotiators; they will bargain for an hour over a cup of coffee, and a day over a camel. But here's the secret the well-judges don't want you to know: the wells have ears. Every cup poured, every promise made at the edge, every drop drawn—the well-spirits are listening. And they have opinions. They favour the generous. They punish the greedy. And they never, ever forget a face.

\vspace{0.5em}

The Fhara have no single king. Their confederacy is a web of marriage alliances, water-sharing treaties, and caravan tariffs. The desert is not empty—it is a ledger of dry wells, abandoned caravanserais, and the bones of those who could not share. A Fhara merchant can cross the waste with only a waterskin and a coffee pot, because every well has a keeper, every keeper a price, and every price can be negotiated. But the negotiation is a ritual. Three cups: the first is welcome, the second is bargaining, the third is a contract. Refuse the third, and you have broken an oath. The well-spirit will remember.

\vspace{0.5em}

Never refuse a date offered by a well-keeper. The fruit is sacred to their well-spirits. Refusing it ends the conversation—and your access to water.

\vspace{1em}

\noindent\textit{— Well-Judge Fatima al-Wadi, who carries a dipper made from a date-palm leaf that rots after each case}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Dates and thin coffee—the dates are from the well-keepers' groves, the coffee is brewed with cardamom and camel milk. The milk is sour. The sourness is for endurance.

\textbf{What does a child learn before they can walk?} To count cups. Fhara children are taught the eight cups of negotiation: Welcome, Bargain, Contract, Witness, Debt, Payment, Forgiveness, and the ninth—the cup that is never poured. The ninth cup is for the well-spirit. The well-spirit is always thirsty.

\textbf{What is the first thing you notice about a stranger?} Their water-skin. A stranger whose water-skin is full has not traveled far. A stranger whose water-skin is dry has traveled too far. A stranger who carries no water-skin is either a well-keeper or a fool. Well-keepers are honoured. Fools are buried.

\textbf{What is the most common crime?} Water theft—but water theft is not a crime, it is a declaration of feud. The well-spirit records every theft in the aquifer. The aquifer is the court. The judge is the well. The verdict is thirst.

\textbf{What does no one talk about but everyone knows?} The wells are not natural. They were carved by the first Fhara, who made a bargain with the sand. The bargain was for water. The price was memory. The Fhara remember everything. They cannot forget. That is the curse of the wells.

\textbf{Where do people go when they die?} To the oasis. The dead are buried in the date groves, their graves marked with a single stone. The stone is not carved. The dead's name is spoken at sunrise. If the name is remembered, the dead rests. If it is forgotten, the dead walks.

\textbf{How do you apologize for a serious wrong?} You offer a cup—the ninth cup, poured from your own water-skin. The cup is placed at the well's edge. If the well-spirit drinks, you are forgiven. If the water evaporates, you are not. If the cup cracks, your debt is doubled.

\textbf{What is the most important festival?} The Filling of the Wells (spring). Every well is cleaned and blessed. The water-clerks recite the names of the year's debtors. The debtors stand at the water's edge. If their reflection wavers, they are forgiven. If it holds, they must pay.

\textbf{What is the secret that could destroy a powerful person?} The water-clerks' ledger is a forgery. The original was lost in a sandstorm three generations ago. The clerks have been inventing water rights ever since. The wells do not care. The wells remember the original. The wells are waiting.
\end{almanac}

\newpage
\input{../generators/fhara.tex}
\newpage

% ------------------------------
% CHAPTER: SIDHI
% ------------------------------
\chapter{Sidhi}
\emph{The Diaspora of the Sealed Coast.}

\noindent
{\Large
\textit{Galanina is not a place on a map. It is a promise that our ancestors made to the sea: we will remember, and we will return. The Covenant gave us courts, but the sea gave us memory. The Admiralty gives us tariffs. We give each other salt and stories.}

\vspace{0.5em}

I am a Freedman Sailor. My grandmother wore a brass collar. My mother wore a freedman's seal. I wear a captain's ring. The slavers took our names, but they could not take the tide. The tide always comes back. So do we.

\vspace{0.5em}

The Sidhi are a people of the coast and the diaspora. Their ancestral homeland is Galanina, a rugged shore of limestone cliffs and ancient ports far to the west. The Ashaani magocracy enslaved them for generations, but Galanina broke free under the Covenant of Mykkiel. Now the Sidhi are scattered, fierce, proud, and never forget a debt. But here's the secret the Admiralty doesn't want you to know: the Sidhi never stopped running. The Long Trail—the forced migration that scattered them across the Amaranthine—is not a tale told to children. It is a silence. A pause in conversation when someone mentions the eastern passes or the southern coast. An old woman who stops mid-sentence, touches her brass key, and changes the subject. The Sidhi do not forget, but they have learned that memory is a kind of weight—and they have carried it far enough.

\vspace{0.5em}

The Sidhi of the Brass Coast run the ropeways, the spice auctions, and the port guilds. Their speech is a creole of Ashaani and coastal tongues, their law is maritime, and their symbol is the lighthouse lamp—a green flame that shows the faces of the drowned, so that the dead may witness the living. They are welcome everywhere and trusted nowhere. They prefer it that way.

\vspace{0.5em}

Never ask about ancestry. The Ashaani slavery raids are a fresh wound. Asking which clan someone belongs to is asking how their grandparents were captured, sold, or killed.

\vspace{1em}

\noindent\textit{— Captain Idris, Sidhi lighthouse commander, whose left eye is a brass lens}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Salted fish and flatbread—the fish is dried on the docks, the bread is baked in communal ovens. The ovens are fuelled by driftwood. The driftwood carries the names of lost ships.

\textbf{What does a child learn before they can walk?} To read the lighthouse codes. Sidhi children are taught the eight lamp-flashes: Safe, Danger, Storm, Pirate, Customs, Quarantine, Home, and the ninth—the flash that signals a freedman's return. The ninth flash is never used. The ninth flash is a promise.

\textbf{What is the first thing you notice about a stranger?} Their seal. Every Sidhi freedman carries a brass seal—the mark of their manumission. A stranger whose seal is worn smooth has been free for generations. A stranger whose seal is sharp has just been freed. A stranger with no seal is not a Sidhi. Not a Sidhi is not a person.

\textbf{What is the most common crime?} Seal-forging—but seal-forging is not a crime, it is a necessity. The Admiralty demands proof of freedom. The proof can be forged. The forgers are Sidhi. The Sidhi do not prosecute forgers. The Sidhi hire them.

\textbf{What does no one talk about but everyone knows?} The brass seals are not brass. They are made from melted chains—the chains that bound the first freedmen. The metal remembers the weight of slavery. The seals are warm. The warmth is the memory of wrists.

\textbf{Where do people go when they die?} To the lighthouse. The dead are not buried. Their ashes are scattered from the tower's top. The wind carries them to the sea. The sea carries them to Galanina. Galanina is the shore of ancestors. The ancestors are waiting.

\textbf{How do you apologize for a serious wrong?} You offer your seal—your own mark of freedom. The seal is melted down and reforged into a chain. The chain is worn by the wronged party. When the wrong is forgiven, the chain is melted again. The metal grows thinner each time. One day, there will be nothing left.

\textbf{What is the most important festival?} The Lighting of the Green Flame (winter solstice). The lighthouse lamps are cleaned and refilled. The green flame is lit. The flame shows the faces of the drowned. The drowned are the ancestors. The ancestors watch. The ancestors do not judge.

\textbf{What is the secret that could destroy a powerful person?} The brass seals are not all equal. Some were forged by the Admiralty itself—trap-seals that mark the wearer as a spy. The spies do not know they are spies. Their reports go to the Admiralty. The Admiralty knows everything. The Admiralty has been playing a very long game.
\end{almanac}

\newpage
\input{../generators/sidhi.tex}
\newpage

% ------------------------------
% CHAPTER: PERESHI
% ------------------------------
\chapter{Pereshi}
\emph{The Roof of the Way.}

\noindent
{\Large
\textit{The fire does not forget. It has witnessed every oath sworn in this valley, every treaty sealed with a dropped coal, every lie told in its light. I do not tend the flame. I listen to what it remembers. And it remembers everything.}

\vspace{0.5em}

I am a Caravan Scribe. My grandfather copied a tax ledger that had not been opened in three centuries. He found a mistake—a single miscounted bale of silk. That mistake started a war. We do not write for the living. We write for the dead who cannot correct their errors.

\vspace{0.5em}

The Pereshi highlands are a vast, windswept plateau, cut by deep river valleys and crowned with snow-capped peaks. Ancient fire temples stand beside caravanserais; terraced vineyards climb the hillsides. The people are scholars, shepherds, and traders—proud, suspicious, and bound by a code of hospitality that demands three days of feasting for any guest, and three generations of vengeance for any insult. But here's the secret the Flame-Keepers don't want you to know: the fire in the temple is not just a flame. It is a witness. It has seen every oath sworn in this valley, every treaty sealed with a dropped coal, every lie told in its light. And it remembers everything. Including the Prince's true name.

\vspace{0.5em}

Once, the Forgotten Ashaani Prince conquered this land. He built garrisons, taxed the passes, and tried to burn every library. When he and his kingdom vanished into the sand—swallowed by the same curse that turned the desert to glass—the Pereshi rose. They threw down his garrisons, brick by brick, and sealed his name in a scroll they have never reopened. They have not forgotten. Every fire temple keeps a coal from the Prince's own brazier—the ninth coal, never lit, a promise that his darkness will not return.

\vspace{0.5em}

Never step on a fallen book. Written words carry the weight of law. To step on a book is to trample a contract, a genealogy, a verdict. The spirits of the words will haunt you.

\vspace{1em}

\noindent\textit{— Flame-Keeper Zarifa, who has never been seen without a rose—fresh or dried}
}

\vspace{1em}

\begin{almanac}
  \textbf{What do people eat for breakfast?} Flatbread with honey and clotted cream—the honey is from highland bees, the cream is from goats that graze the temple steps. The goats are sacred. The goats are also delicious. The bread is baked in ovens heated with coals from the temple fire; families take turns carrying the embers home. A coal that goes out before dawn is a bad omen.
  
  \textbf{What does a child learn before they can walk?} To read a contract. Pereshi children are taught the eight clauses of the Covenant: Offer, Acceptance, Witness, Consideration, Term, Breach, Remedy, and the ninth—the silence where the unspoken terms live. The ninth clause is never written. The ninth clause is the one that kills you. Children also learn to check the ninth coal in the family shrine each morning. If it has grown warm overnight, they must wake the Flame-Keeper.
  
  \textbf{What is the first thing you notice about a stranger?} Their book. Every Pereshi carries a ledger—their personal account of debts owed and collected. A stranger whose ledger is thick has many debts. A stranger whose ledger is thin has many creditors. A stranger with no ledger is a liar. Liars are not fed. But these days, the first thing many notice is whether the stranger casts a shadow. The Prince's shadow has been seen walking alone.
  
  \textbf{What is the most common crime?} Contract forgery—but forgery is not a crime, it is an art. The Pereshi courts are full of forged documents. The judges are forgers themselves. The law is a negotiation. The negotiation never ends. The newest crime is shadow-theft: someone steals another's shadow to feed to the Prince. The victim becomes pale, listless, and cannot lie—but also cannot keep a secret.
  
  \textbf{What does no one talk about but everyone knows?} The Forgotten Prince did not vanish. He is buried beneath the largest fire temple, in a lead-lined coffin sealed with nine locks. The keys are held by the nine Flame-Keepers. The Flame-Keepers have never met. The Flame-Keepers do not know each other's names. The Prince's name is written on the inside of the coffin lid. No one has read it in nine centuries—but lately the lid has begun to sweat. A rust-coloured liquid seeps from the seams.
  
  \textbf{Where do people go when they die?} To the fire. The dead are cremated on the temple pyres. Their ashes are mixed with clay and pressed into bricks. The bricks are used to build new temples. The dead become walls. The walls witness. The walls do not forget—and they whisper the names of those who still live. Some walls have started to whisper a name that is not on any ledger.
  
  \textbf{How do you apologize for a serious wrong?} You offer a coal—a burning ember from the temple fire. The coal is placed in the wronged party's hearth. If it stays lit, you are forgiven. If it smoulders, you are watched. If it goes out, you are cursed. The curse is the fire's silence. The fire will not speak for you again. Some say that if the coal goes out, the Prince's shadow will visit you that night. It asks only one question: “Why should I not take your name?”
  
  \textbf{What is the most important festival?} The Reading of the Ledgers (autumn). Every household brings its ledger to the temple. The ledgers are read aloud. The debts are tallied. The dead are named. The fire listens. The fire does not interrupt—but the ninth coal flickers when a lie is spoken. This year, the coal has not stopped flickering.
  
  \textbf{What is the secret that could destroy a powerful person?} The ninth coal is not coal. It is a piece of the Forgotten Prince's heart, petrified by the curse that swallowed his kingdom. The Flame-Keepers do not know. They think it is a relic. The heart is still warm. The heart is still beating. The heart is waiting to be reunited with the rest of him. And someone has been feeding it blood—a drop each night, from a different Flame-Keeper. No one knows who. No one wants to know who.
  \end{almanac}

\newpage
\input{../generators/pereshi.tex}
\newpage

% ------------------------------
% CHAPTER: KUVANI
% ------------------------------
\chapter{Kuvani}
\emph{The Remount Steppe.}

\noindent
{\Large
\textit{The grass does not belong to us. We belong to the grass. It tells us where to ride, when to move, how long to stay. A clan that forgets the grass's song is a clan that will be buried by it. I do not sing to the wind. I listen. The wind sings through me.}

\vspace{0.5em}

I am a Remount Master. A horse's name is its lineage. Its lineage is its debt. You do not ride a Kuvani horse. You borrow it. And when you return it, you will tell me a story that the horse has not already heard. If you lie, the horse will throw you next time.

\vspace{0.5em}

Horses, grass, and wind-songs—the Kuvani don't just pay in stories; they auction the right to the next remount. They range across the eastern steppe, a vast grassland east of the Valewood. Their horsemanship and remount stations make them indispensable to any caravan that crosses the open plain. The Kuvani have no cities; their courts are circles of saddles, and their laws are sung. But here's the secret the Wind-Singers don't want you to know: the sky is not empty. Kuva, the sky father, sees everything—every oath sworn in the open, every horse born, every blade of grass bent by the wind. But Kuva does not speak. He listens. And he is patient. The Kuvani have no temples because the sky is their shrine. Their dead are not buried; they are given to the sky on funeral platforms, that Kuva may carry their souls to the ancestral herd that rides the northern lights.

\vspace{0.5em}

A horse is not a mount—it is a relative. Its name is a lineage stretching back to the first mare that Kuva breathed life into. A Kuvani who cannot name his horse's ancestors is a Kuvani without a shadow. The steppe is not empty; it is full of ghosts: the ghost herd of the salt pan, the sky-road markers that hum at dusk, the Iron Corral where blades whisper the names of the dead.

\vspace{0.5em}

Never ride a Kuvani horse without asking its name. The horse will answer—by throwing you. If it throws you, you have insulted its lineage.

\vspace{1em}

\noindent\textit{— Alp Arslan, Remount Master, who has never been seen sleeping}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Fermented mare's milk and dried meat—the milk is from the herd, the meat is from the hunt. The hunt is a ceremony. The ceremony is a debt to the grass.

\textbf{What does a child learn before they can walk?} To ride. Kuvani children are placed on a horse before they can stand. The horse is patient. The horse is old. The horse remembers the child's grandfather. The grandfather taught the horse to ride.

\textbf{What is the first thing you notice about a stranger?} Their saddle. A stranger whose saddle is worn has ridden many miles. A stranger whose saddle is new has many miles ahead. A stranger who carries no saddle is a ghost. Ghosts do not need saddles.

\textbf{What is the most common crime?} Horse theft—but horse theft is not a crime, it is a declaration of feud. The horse is family. A stolen horse is a stolen relative. The feud lasts until the horse is returned or the thief's line is ended. The feud never ends.

\textbf{What does no one talk about but everyone knows?} The ghost herd is not a legend. It is real—a herd of spectral horses that rides the salt pan on moonless nights. The horses are the souls of riders who died without honour. They ride forever, searching for a battle that will never come.

\textbf{Where do people go when they die?} To the sky. The dead are placed on funeral platforms, wrapped in felt. The birds take the flesh. The wind takes the bones. The sky takes the soul. The soul joins the ancestral herd. The herd rides the northern lights.

\textbf{How do you apologize for a serious wrong?} You offer a horse—the best horse in your herd. The horse is given to the wronged party. If they accept, the debt is paid. If they refuse, the debt is doubled. If they kill the horse, the debt is blood. Blood is never repaid.

\textbf{What is the most important festival?} The Ride of the Ancestors (autumn equinox). The clans ride to the salt pan. The ghost herd appears at midnight. The living ride with the dead. The dead remember their names. The living remember theirs.

\textbf{What is the secret that could destroy a powerful person?} The Khagan's horse is a ghost. It was killed in battle twenty years ago. The Khagan does not know. He rides a skeleton wrapped in hide. The horse does not tire. The horse does not eat. The horse does not forgive.
\end{almanac}

\newpage
\input{../generators/kuvani.tex}
\newpage

% ------------------------------
% CHAPTER: TULKANI
% ------------------------------
\chapter{Tulkani}
\emph{The Itinerant Courts.}

\noindent
{\Large
\textit{The road does not forgive a debt. It only reschedules it. A promise spoken under the open sky is witnessed by the wind, the grass, and the wheels of every cart that passes. I do not write judgments. I tie knots. The knots remember. When you are ready to pay, untie one. If you are not ready, tie another.}

\vspace{0.5em}

I am a Wagon-Keeper. My wagon is my hearth, my court, my grave. The Hollow cannot cross the circle of wheels unless invited. The tax collector cannot enter without a story. The lord cannot demand rent unless we camp on his land for nine nights. On the eighth night, we leave. That is the law.

\vspace{0.5em}

The Tulkani have no fixed homeland; they are wanderers who follow the seasons and the trade routes. Their wagons are their hearths, their circles are their courts, and their knots are their ledgers. They are welcome everywhere and trusted nowhere—a position they wear with wry humour. They spread the cult of Ikasha—She Who Sleeps—not through temples, but through dream-cords tied in the dark. But here's the secret the Caravan Judges don't want you to know: the knots are not just records. They are alive. A knot tied in false witness will tighten until blood is drawn. A cord cut in anger will become a ghost that follows the oath-breaker until the promise is fulfilled. And the ninth knot—the one that is never tied—is the space where the Hollow waits.

\vspace{0.5em}

The Tulkani have no king, no judges, no written law—only the Caravan Circle, where disputes are settled by storytelling. A liar is exposed not by evidence but by the knots on his cord. Ikasha's cult is not a religion; it is a survival tactic—the dream inside the wagon, the whisper in the creaking wheel. The Tulkani pay no permanent taxes, but they pay tolls in stories. A lord who demands coin will be told a story so boring that his crops fail. A lord who offers hospitality will be told a story so beautiful that his harvest doubles. The Tulkani do not track debts; they track reciprocity. What you give, you get back—maybe tomorrow, maybe in the next valley. But the circle remembers.

\vspace{0.5em}

Never sleep outside the circle at night. The wagon ring is sanctuary. Outside it, the Hollow walks. A Tulkani will not rescue you—they will only mourn you at dawn.

\vspace{1em}

\noindent\textit{— Elder Sarnai, Caravan Judge of the Nine Knots, whose cord drags on the ground}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Tea and hardtack—the tea is brewed from leaves gathered along the road, the hardtack is baked at the start of the season and lasts until the snow. The hardtack is always stale. The tea is always fresh. The contradiction is a lesson.

\textbf{What does a child learn before they can walk?} To tie a knot. Tulkani children are taught the eight journey-knots: Departure, Arrival, Promise, Debt, Payment, Witness, Apology, and the ninth—the knot that is never tied. The ninth knot is for the Hollow. The Hollow is not a creditor. The Hollow is a reminder.

\textbf{What is the first thing you notice about a stranger?} Their cord. Every Tulkani carries a knotted cord—their life's ledger. A stranger whose cord has many knots has traveled far. A stranger whose cord has few knots has just begun. A stranger with no cord is a ghost. Ghosts do not travel. Ghosts wait.

\textbf{What is the most common crime?} Story-theft—but story-theft is not a crime, it is a compliment. A stolen story is a story worth telling. The original teller is honoured. The thief is welcomed. The story grows with each telling. The story never dies.

\textbf{What does no one talk about but everyone knows?} The Tulkani are not wanderers by choice. They were exiled from their homeland nine centuries ago, for a crime that no one remembers. The crime is written in the ninth knot of the Elder's cord. The Elder's cord is too long to read. The Elder's cord drags on the ground.

\textbf{Where do people go when they die?} To the road. The dead are not buried. Their bodies are left at a crossroads. The road takes them. The road carries them to the next valley, the next camp, the next telling. The dead become stories. Stories are immortal.

\textbf{How do you apologize for a serious wrong?} You untie a knot—one knot from your cord, chosen by the wronged party. The knot is burned. The ash is scattered on the wind. If the wind blows toward the wronged party, you are forgiven. If it blows away, you are not. If it is still, the Hollow has accepted your apology. The Hollow's acceptance is a warning.

\textbf{What is the most important festival?} The Tying of the New Cord (new year). Every Tulkani receives a new cord, unknotted. The old cord is burned. The ash is mixed with water and drunk. The drink tastes of memory. The memory is the year that has passed. The year is never forgotten.

\textbf{What is the secret that could destroy a powerful person?} The Tulkani are not wanderers. They are pilgrims—walking toward a destination that no longer exists. Their homeland was swallowed by the desert. The desert is still hungry. The Tulkani are still walking. They will never arrive.
\end{almanac}

\newpage
\input{../generators/tulkani.tex}
\newpage

% ============================================================
% PART VI: DHAHARA — THE KINGDOM OF BRASS
% ============================================================
\part{Dhahara — The Kingdom of Brass}
% After \part{Dhahara — The Kingdom of Brass}
\begin{quote}
  \small
  \textit{``The mountain weighs every breath, every ingot, every oath. Pay in iron, pay in silence, or pay in blood. And when the ninth lease cracks, even the elephants will remember who fed them during the famine.''}
  \noindent — Dusana of the Raven Road, at the Viceroy's Hall in Sarvistan
  \end{quote}
% ------------------------------
% CHAPTER: THE BRASS COAST (REVISED)
% ------------------------------
\chapter{The Brass Coast}
\emph{Shore of Silver and Spice.}

\noindent
{\Large
\textit{“The tariff is the law. The law is the tariff. I do not care if your goods are stolen, blessed, or weeping. I care that you pay. The Admiralty does not rule this coast. The ledger rules. I am merely its pen.”}

\vspace{0.5em}

I am a spice merchant, grandson of a raja who watched the first Kahfagian customs house rise on our beach. The Admiralty calls me a subject. The Sidhi call me a partner. I call myself a son of the Brass Coast. My grandfather’s sword hangs above my desk. One day, I may need it again. The sword has nine notches. Each notch is a promise the Admiralty broke.

\vspace{0.5em}

The Brass Coast is Dhahara’s southern rim: a humid, feverish coastline of palm groves, salt pans, and crowded port cities. The Kahfagian Admiralty administers it as a “client province,” but three powers bleed into each other here, and none can draw a clean breath. The Admiralty holds the ports and the tariff ledgers. The Sidhi diaspora runs the ropeways, the lighthouse courts, and the freedman’s seal. And the native Dhaharan rajas — the hill lords who have lived here since before any foreign flag flew — grow the pepper, mine the iron, and watch the coast from inland fortresses, waiting.

\vspace{0.5em}

Here, a man can buy anything — a ship, a spice lot, a freedman’s seal — and sell anything except his own shadow. The Admiralty’s Maritime Court settles disputes with a judge’s gavel and a marine’s cutlass. The Sidhi lighthouse courts offer an alternative: justice by green flame, witnessed by the drowned. And the rajas enforce their own law in the hills, where the Admiralty’s writ does not run. But here’s the secret the Admiralty doesn’t want you to know: the ropeways that carry spice baskets from ship to warehouse are old, and the dead still pull their weight — not just Sidhi dead, but the drowned of every nation who could not pay their passage. The coast remembers them all. The ninth rope on every ropeway is always frayed. No one replaces it. That rope carries the debt.

\vspace{0.5em}

\textbf{Never board a ship that has a green lantern at the mast.} It means the captain is under a death vow — and you are the payment. If the lantern burns blue, run. The blue flame is the Admiral’s oracle, and she does not make bargains.

\vspace{1em}

\noindent\textit{— Raja Govind of the Kalinga foothills, who has never seen the sea and refuses to start}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice and fermented fish paste — the rice is grown in coastal paddies, the fish is caught by the lighthouse keepers’ children. The children are paid in copper. The copper is melted into seals. The seals are sold to whoever can afford them. The ninth child never takes payment. No one knows why.

\textbf{What does a child learn before they can walk?} To read a lighthouse code. Brass Coast children are taught the eight lamp-flashes: Tariff, Inspection, Quarantine, Smuggler, Pirate, Rescue, Home, and the ninth — the flash that means “the collector is coming.” The ninth flash is never used. The ninth flash is a warning. The children learn it anyway.

\textbf{What is the first thing you notice about a stranger?} Their papers. The Admiralty requires a manifest for every shipment, a seal for every trader, a stamp for every dock. A stranger whose papers are in order is a merchant. A stranger whose papers are missing is a smuggler. A stranger whose papers are too perfect is a spy. The spies are Admiralty. The Admiralty spies on itself. The ninth page of every manifest is always blank. That is where the dead travel.

\textbf{What is the most common crime?} Tariff evasion — but tariff evasion is not a crime, it is a profession. The tariffs are set by the Admiralty. The Admiralty is corrupt. The evaders pay the Admiralty directly. The payment is called a “consulting fee.” The fee is tax-deductible. The ninth fee is always paid in salt, not silver. The salt comes from the Freedman’s Grave.

\textbf{What does no one talk about but everyone knows?} The lighthouse lamps are not fuelled by oil. They are fuelled by the fat of the drowned — whales, sailors, slaves who did not survive the crossing. The fat burns green. The green is the colour of memory. The memory is the dead’s. The ninth lamp on every tower is never lit. It is reserved for the collector.

\textbf{Where do people go when they die?} To the sea. The dead are weighted with brass and dropped in the deep channel. The channel is called the Freedman’s Grave, but it holds rajas and pirates alike. The grave is unmarked. The grave is the entire coast. The ninth body dropped each year is never weighted. It floats. It always drifts toward the Admiral’s palace.

\textbf{How do you apologize for a serious wrong?} You offer a tariff — a payment of silver, calculated by the wronged party’s estimate of their loss. The tariff is paid to the Maritime Court. The court takes a percentage. The percentage is the judge’s fee. The judge’s fee is negotiable. The ninth tariff ever collected was paid in a memory. The judge still remembers it. He wishes he did not.

\textbf{What is the most important festival?} The Burning of the Ledgers (new year). Every merchant’s ledger is brought to the lighthouse. The ledgers are read aloud. The debts are tallied. The ledgers are burned. The ash falls on the water. The water carries the debts to the sea. The sea does not forgive. The sea reschedules. The ninth ledger is never burned. It is kept in a brass box at the bottom of the Freedman’s Grave. It contains the Admiralty’s own debts.

\textbf{What is the secret that could destroy a powerful person?} The lighthouse lamps are not all green. One lamp burns blue — the Admiral’s personal beacon. The blue flame is not fat. It is a spirit — a Sidhi oracle who was bound to the lighthouse nine generations ago. She sees everything. She speaks to no one. The Admiral does not know she is still alive. Her ninth prophecy was never recorded. It said: “The collector will come by sea, and the ledger will burn from the bottom up.”
\end{almanac}

\newpage
\input{../generators/brass_coast.tex}
\newpage

% ------------------------------
% CHAPTER: THE NINE KINGDOMS (REVISED)
% ------------------------------
\chapter{The Nine Kingdoms}
\emph{The Peacock Throne and Its Rivals.}

\noindent
{\Large
\textit{“The Peacock Throne is not a seat. It is a promise made by the first sun king: that the nine kingdoms would remain nine, never one, never fewer. I do not rule. I balance. A peacock's tail has nine eyes. Each eye watches a different kingdom. When one closes, the throne weeps.”}

\vspace{0.5em}

I am a Temple Dancer. My dance is not for the king. It is for the monsoon. If my ankle bells ring true, the rains come. If they falter, the wells dry. They have never faltered. I do not sleep before a performance. I dream the rhythm. And I dream of the missing ninth.

\vspace{0.5em}

Between the Satrapies plateau and the Brass Coast lies a fertile, contested lowland: the land of the Nine Kingdoms. This is the heart of Dhahara, where ancient dynasties still claim descent from the sun, and where the Peacock Throne of Ahimsa casts its shadow over half a dozen rival courts. Here, rivers are borders, elephants are war machines, and every harvest is a negotiation between kings who have never met but whose grandfathers swore blood oaths.

But here's the secret the Maharaja doesn't want you to know: the number “nine” is a lie. There have never been nine kingdoms. The number is a ritual — a prayer that the Hollow will not notice that one of the nine is always missing. The throne weeps because the missing kingdom is always the one that mattered. Its debt is still compounding.

\vspace{0.5em}

The Covenant has a strong presence in the sacred cities; the Aeler have counting houses in the iron towns; the Sidhi sail up the rivers to trade. But the Nine Kingdoms belong to the Dhaharan — to their poets, their temple dancers, their bandit kings, and their queens who rule from behind silk screens. A king's authority rests on his genealogy — how many generations back to the sun, how many treaties sealed, how many marriages made. But a king’s true power rests on how much of his family’s debt has been paid. The Peacock Throne of Ahimsa is the nominal centre, but the other eight kings vie for influence, and the balance shifts with every monsoon.

\vspace{0.5em}

\textbf{Never praise a king's elephant.} The elephant is sacred; your praise will be taken as an insult to the king's own mount. Compliment the mahout instead. The mahout will tell the elephant. The ninth elephant in the royal stables has no mahout. It waits.

\vspace{1em}

\noindent\textit{— Temple Dancer Mira, who has stopped three droughts and started none}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with mango pickle — the rice is from the royal granaries, the mangoes are from the king's orchards. The king's orchards are irrigated by the river. The river is contested. The contest is a negotiation. The negotiation is breakfast. The ninth grain of rice in every bowl is left uneaten. It is an offering to the missing kingdom.

\textbf{What does a child learn before they can walk?} To recite their genealogy. Nine Kingdoms children are taught the names of their ancestors back to the sun. A child who cannot recite is a child without a kingdom. A child without a kingdom is a servant. Servants are not fed. The ninth ancestor is never named. That is the gap where the Hollow sits.

\textbf{What is the first thing you notice about a stranger?} Their caste mark. The Nine Kingdoms have nine castes — priest, warrior, merchant, farmer, artisan, servant, outcast, foreigner, and the ninth — the caste that has no name. The ninth caste is the king’s spies. The spies are everywhere. The spies have no caste mark. Their faces are always the same. No one can describe them.

\textbf{What is the most common crime?} Tax evasion — but tax evasion is not a crime, it is a political statement. The king's tax collectors are also the king's spies. A merchant who evades taxes is a merchant who has something to hide. The king wants to know what. The king will find out. The ninth tax collector vanished last year. His shadow still walks the counting house.

\textbf{What does no one talk about but everyone knows?} The Peacock Throne is empty. The Maharaja died seven years ago. His death was concealed by the courtiers, who rule in his name. The courtiers are the nine ministers. The ninth minister is a eunuch who controls the palace guards. The eunuch has the real power. He has no shadow. The guards do not notice.

\textbf{Where do people go when they die?} To the river. The dead are cremated on the ghats. Their ashes are scattered on the water. The water flows to the sea. The sea is the kingdom's border. The border is a negotiation. The negotiation never ends. The ninth ghat is reserved for the missing kingdom. It has never been used. It is always warm.

\textbf{How do you apologize for a serious wrong?} You offer a dance — a performance at the temple of the sun. The dance is judged by the priest. If the priest smiles, you are forgiven. If the priest weeps, you are not. If the priest is silent, you are cursed. The curse is the sun's glare. The sun will not look away. The ninth dance ever performed was for a debt to the Hollow. The dancer is still dancing. No one watches.

\textbf{What is the most important festival?} The Coronation of the Monsoon (summer). The king offers a golden elephant to the rain clouds. The clouds accept or refuse. The clouds have refused for nine years. The king's treasury is empty. The king's patience is not. The ninth golden elephant was hollow. It was filled with ashes. The ashes belonged to the ninth king.

\textbf{What is the secret that could destroy a powerful person?} The golden elephant is hollow. It is filled with the ashes of the ninth king — the one who tried to unite the kingdoms and failed. His ghost rides the monsoon winds. His ghost is still trying. His ghost will never succeed. But every year, the clouds refuse because they are waiting for him to ask. He never does.
\end{almanac}

\newpage
\input{../generators/nine_kingdoms.tex}
\newpage

% ------------------------------
% CHAPTER: THE HIMADRI FOOTHILLS
% ------------------------------
\chapter{The Himadri Foothills}
\emph{The Iron Passes (Aeler Focus).}

\noindent
{\Large
\textit{The mountain does not speak in words. It speaks in the crack of a bearing, the echo of a hammer, the taste of ore on the tongue. We Aeler do not conquer stone. We negotiate with it. The lease is not a document. It is the weight of the keystone that holds the roof.}

\vspace{0.5em}

I am a Mine-Breath. I carry a bell that counts my breaths. Eight is safe. Nine is the mountain's. My grandfather counted to nine once. The mountain took his shadow. He lived another thirty years, but he never cast a shadow again.

\vspace{0.5em}

The Ninth Seal mine isn't just holding a demon—it's a seal market, and oh, how the Vent-Prior loves a good deal on demon containment. The Aeler do not rule through armies; they rule through the maintenance contracts that keep the mines from collapsing, the bridges from falling, and the passes from freezing. A vent-prior's authority rests on his ability to read the mountain's cracks; a mine-captain's on her skill with a tuning fork. But here's the secret the Vent-Prior doesn't want you to know: the demon in the Ninth Seal is not the only thing the mountain is holding. There are older things deeper down. The Aeler sealed them first. The Ninth Seal is just the newest.

\vspace{0.5em}

The elephant caravans haul ore down to the Brass Coast; mule trains carry silver and tea up to the mountain holds. The Aeler engineers speak a creole of their own tongue; the Dhaharan nobles wear Aeler breath-counters as jewellery; and deep beneath the oldest mine, a demon sleeps behind a Covenant seal—a fragment of the thing that turned the Red Desert to glass. The seal is cracking. The demon whispers through the cracks. It offers favours. It asks for small betrayals. It is very patient.

\vspace{0.5em}

Never count your breaths aloud in a mine. The mountain listens, and if it hears you reach the ninth, it will take one—a breath, a memory, a shadow.

\vspace{1em}

\noindent\textit{— Mine-Breath, who carries her grandfather's bell and has never reached nine}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Iron rations and tea—the rations are hardtack and dried meat, the tea is brewed from leaves that grow only above the snow line. The leaves are harvested by the Vent-Prior's apprentices. The apprentices are chosen by lottery. The lottery is a death sentence.

\textbf{What does a child learn before they can walk?} To read a pressure gauge. Himadri children are taught the eight pressures: Safe, Warning, Danger, Collapse, Flood, Fire, Seal, and the ninth—the pressure that means the mountain is breathing. The mountain's breath is the miners' death.

\textbf{What is the first thing you notice about a stranger?} Their bell. Every miner carries a breath-counter bell. A stranger whose bell rings too fast has been running. A stranger whose bell is silent is dead. A stranger whose bell rings the ninth note is a demon. The demons are sealed. The seals are cracking.

\textbf{What is the most common crime?} Ore smuggling—but ore smuggling is not a crime, it is a perquisite. The mine-captains take a cut of every shipment. The cut is called "the Vent-Prior's fee." The fee is negotiable. The negotiation is a ritual. The ritual involves tea and threats.

\textbf{What does no one talk about but everyone knows?} The Ninth Seal is not a seal. It is a door—a gate to a prison that holds something that should not be named. The seal's keepers do not know what is inside. The keepers do not want to know. The keepers are paid to not know.

\textbf{Where do people go when they die?} To the mountain. The dead are not buried. Their bodies are left in worked-out galleries. The mountain absorbs them. The mountain's memory grows. The mountain remembers every breath it has taken.

\textbf{How do you apologize for a serious wrong?} You offer a breath—one of your own, counted and witnessed. The breath is sealed in a brass capsule and placed in the Vent-Prior's ledger. The ledger is the court. The court's judgment is the mountain's. The mountain's judgment is final.

\textbf{What is the most important festival?} The Counting of the Breath (winter solstice). Every miner's breath-counter is read aloud. The miners with the lowest counts are promoted. The miners with the highest counts are retired. Retirement is a sealed gallery. The gallery is a tomb.

\textbf{What is the secret that could destroy a powerful person?} The demon in the Ninth Seal is not a demon. It is an Aeler—the first Vent-Prior, who volunteered to be sealed with the mountain's hunger. He is still alive. He has been counting his breaths for nine centuries. He has reached the ninth count. He is still counting.
\end{almanac}

\newpage
\input{../generators/himadri_foothills.tex}
\newpage

% ============================================================
% PART VII: THE EASTERN REALMS
% ============================================================
\part{The Eastern Realms}
% After \part{The Eastern Realms}
\begin{quote}
  \small
  \textit{``The maps of the empire draw the East as a single sweep – Sihai, Nihori, Ayokha – all pressed together like tiles on a game board. The roads know better. Sihai measures time in dynasties; Nihori counts it in storms. Ayokha negotiates between them with the patience of the monsoon. And beneath them all, the older peoples keep their own thresholds, their own debts, their own ways of not being seen.''}
  \noindent — Dusana of the Raven Road, at the edge of the Inland Sea
  \end{quote}
% ------------------------------
% CHAPTER: SIHAI
% ------------------------------
\chapter{Sihai}
\emph{The Central Kingdom.}

\noindent
{\Large
\textit{The Mandate of Heaven is not a right. It is a loan. The emperor borrows legitimacy from the ancestors, from the land, from the Hollow itself. When the loan is due, the empire falls. I do not serve the emperor. I serve the ledger that tracks the interest.}

\vspace{0.5em}

I am a Provincial Censor. My brush is heavier than any sword. A single stroke can erase a name, a family, a dynasty. The emperor reads my reports and trembles. He knows that I have written nine reports this year. The ninth is about him.

\vspace{0.5em}

Sihai is the central kingdom, the seat of the Emperor and the Mandate of Heaven. Its canals carry grain from the south; its examination halls produce bureaucrats who govern by the brush. Power here is measured in ink, not iron. The Censorate watches everyone; the Salt Monopoly controls trade; the Lethai-thora keep the secrets that the Mandate prefers to forget. But here's the secret the Emperor doesn't want you to know: the Mandate is not a divine right. It is a loan. The ancestors are the creditors. The Hollow collects the interest. Every dynasty that has ever fallen did so because the interest came due and the emperor could not pay.

\vspace{0.5em}

Beneath the mortal bureaucracy runs a deeper one: the Celestial Bureaucracy of gods, ghosts, and ancestors, where every official has a heavenly counterpart and every erased name becomes a hungry ghost. The Lethai-thora—the ancient archivists who taught the first emperors to write—still keep the hidden ledgers, the ones that record what the Censorate burns. A well-written couplet can appease a ghost; a forged seal can summon a demon. The Hollow's ninth debt is the unpaid interest on the Mandate itself.

\vspace{0.5em}

Never speak of succession before naming the Emperor. The Censorate has ears in every tea house. A misplaced word can erase your name from the rolls.

\vspace{1em}

\noindent\textit{— Quiet Chancellor Wei, who has never smiled and whose seal is a laughing mouth}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Congee with preserved vegetables—the congee is rice boiled until it is memory, the vegetables are salted and aged in clay jars. The jars are sealed with the emperor's own wax. The wax is a promise. The promise is that the vegetables will not poison you. The promise is sometimes broken.

\textbf{What does a child learn before they can walk?} To write their name. Sihai children are taught the eight strokes of the character "person"—the ninth stroke is the one that turns a name into a signature. A signature is a contract. A contract is a debt. A debt is a life.

\textbf{What is the first thing you notice about a stranger?} Their chop—the seal they carry to sign documents. A stranger whose chop is jade is a noble. A stranger whose chop is brass is a clerk. A stranger whose chop is wood is a merchant. A stranger with no chop is a ghost. Ghosts cannot sign contracts. Ghosts cannot be trusted.

\textbf{What is the most common crime?} Examination fraud—but fraud is not a crime, it is a family business. The examinations determine who governs. The questions are leaked in advance. The answers are memorized. The examiners are bribed. The system is a sieve. The sieve is called "merit."

\textbf{What does no one talk about but everyone knows?} The Emperor is a figurehead. The real power is the Censorate—nine ministers who have never been seen together. Their names are not recorded. Their faces are not painted. Their orders are whispered. The Emperor signs what they tell him to sign. The Emperor has not refused in forty years.

\textbf{Where do people go when they die?} To the ancestral tablets. The dead are not buried. Their names are written on wooden tablets, placed in the family shrine. The tablets are fed with rice and tea. The dead eat the steam. The dead are never full.

\textbf{How do you apologize for a serious wrong?} You offer a calligraphy—a scroll written in your own blood. The scroll is presented to the wronged party's ancestral shrine. If the ancestors accept, you are forgiven. If the scroll bursts into flame, you are not. If the scroll writes itself, the ancestors are angry. The ancestors' anger is a famine.

\textbf{What is the most important festival?} The Festival of the Ancestors (autumn). The tablets are cleaned and polished. The dead are invited to feast. The feast is laid out on the family altar. The food is removed after an hour. The dead have eaten the essence. The essence is memory.

\textbf{What is the secret that could destroy a powerful person?} The Censorate's ninth minister is a woman—the emperor's own mother, who was supposed to have died in childbirth. She faked her death to escape the politics of the court. She has been running the Censorate from a hidden villa in the western hills. The emperor does not know. The emperor's mother signs his death warrants. He signs them too.
\end{almanac}

\newpage
\input{../generators/sihai.tex}
\newpage

% ------------------------------
% CHAPTER: NIHORI
% ------------------------------
\chapter{Nihori}
\emph{The Isles of the Dawn Spirit.}

\noindent
{\Large
\textit{I ruled the waves before the Hollow learned to walk. My empire sank because I spoke a name I should not have. Do not speak my name. Do not ask what it was. The sea remembers. That is enough.}

\vspace{0.5em}

I am a Wanderer. I carry a broken sword and a grudge. The sword was my father's. The grudge is mine. The man who killed him died nine years ago. I still walk the isles, looking for his ghost. The ghost is looking for me too.

\vspace{0.5em}

Nihori is a chain of volcanic islands, shrouded in mist and lashed by storms. Here, the old empire sank beneath the waves, and the Lethai-al guard the drowned kingdom's memory. The people are wanderers, duelists, and shrine maidens. They worship the storm spirits and fear the ninth wave. The Hollow breathes in the salt foam; the Oath-Keepers ride the tide. But here's the secret the Shogun doesn't want you to know: the old empire did not sink. It was pulled under. And whatever pulled it is still down there, waiting for the ninth wave to break the right way.

\vspace{0.5em}

The Shogun sits in a half-ruined castle, his authority challenged by three rival daimyo who sharpen their blades and their treaties. Oni cults whisper in the ash wastes, offering forgotten bargains. The Storm Emperor's ghost does not rule—but his silence haunts every island. The ninth taboo is absolute: never speak the name of the drowned capital. The sea will hear you, and it will ask why you called. And when the sea asks a question, it expects an answer—in blood, in memory, or in a name you did not mean to give.

\vspace{0.5em}

Never speak the name of the drowned capital. The sea will hear you, and it will ask why you called.

\vspace{1em}

\noindent\textit{— The Storm Emperor's Ghost, who appears as a column of steam over the water}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Grilled fish and pickled vegetables—the fish is caught in the caldera lakes, the vegetables are fermented in clay pots buried in volcanic ash. The ash is warm. The warmth is the mountain's breath. The mountain is sleeping. The mountain dreams of the drowned empire.

\textbf{What does a child learn before they can walk?} To read the tide. Nihori children are taught the eight wave-heights: Calm, Ripple, Swell, Storm, Tsunami, and the ninth—the wave that has no height because it is the sea itself rising. The ninth wave is the drowning. The drowning is the empire's memory.

\textbf{What is the first thing you notice about a stranger?} Their sword. Nihori warriors carry their ancestors' blades, passed down for generations. A stranger whose sword is rusted has forgotten their lineage. A stranger whose sword is sharp has remembered. A stranger with no sword is a coward. Cowards are not fed.

\textbf{What is the most common crime?} Dueling—but dueling is not a crime, it is a conversation. The dueling codes are written in verse. The verses are interpreted by the Shogun's court. The interpretations are contradictory. The contradictions are settled by dueling. The circle is beautiful.

\textbf{What does no one talk about but everyone knows?} The drowned capital is not underwater. It is beneath—beneath the caldera, beneath the bedrock, beneath the world. The sea is a lid. The lid is cracked. Something is leaking through. The leaks are the oni. The oni are the empire's regrets.

\textbf{Where do people go when they die?} To the sea cliffs. The dead are placed in caves overlooking the water. The waves sing to them. The songs are the empire's lullabies. The dead listen. The dead do not sleep.

\textbf{How do you apologize for a serious wrong?} You offer a duel—a matched contest with the wronged party's choice of weapon. If you lose, you are forgiven. If you win, you are not. If you refuse, you are exiled. Exile is the open sea. The sea is patient. The sea will wait.

\textbf{What is the most important festival?} The Calling of the Tides (spring). The shrine maidens walk into the water, singing the names of the drowned. The waves answer. The waves speak the names of the living who will die in the coming year. The names are never forgotten. The names are never repeated.

\textbf{What is the secret that could destroy a powerful person?} The Shogun is a ghost. He died in the last oni war, but his body was never found. The daimyo rule in his name because they are afraid of what will happen if they admit he is gone. The oni know. The oni are waiting for the admission. The admission will be the war.
\end{almanac}

\newpage
\input{../generators/nihori.tex}
\newpage

% ------------------------------
% CHAPTER: AYOKHA
% ------------------------------
\chapter{Ayokha}
\emph{The Monsoon Throne.}

\noindent
{\Large
\textit{The monsoon does not come because we pray. It comes because the stars align. I read the stars. I align them. Without me, the rains would fail, the rice would burn, and the kingdom would fall. I am not a servant of the God-King. I am his monsoon.}

\vspace{0.5em}

I am a River Merchant. My boat has more secrets than cargo. The lower deck is for silk. The upper deck is for lies. The tiller is for bribes. I have carried nine spies this year. The ninth was the God-King's own daughter.

\vspace{0.5em}

Ayokha is the land of the Monsoon Throne, where the God-King rules by the rhythm of the rains. Rivers flood, rice terraces glow green, and the wharves are crowded with ships from Dhahara, Sihai, and distant isles. The Lethai-ar witness every contract; the spirit-mediums speak for the nats of the ford. Power here is measured in monsoon schedules, silk vouchers, and the favour of the God-King's astrologer. But here's the secret the astrologer doesn't want you to know: the monsoon does not come because he aligns the stars. The monsoon comes because the river-spirits allow it. And the river-spirits are angry.

\vspace{0.5em}

The nats—hungry spirits who live in eddies and under bridges—demand offerings of betel leaf, candle wax, and forgotten promises. A merchant who ignores a nat's demand will find his boat sunk and his name struck from the ledgers. A farmer who disrespects the monsoon will find his fields barren and his children dreaming of drowning. In Ayokha, the rain is not weather—it is a creditor. And the God-King is behind on his payments.

\vspace{0.5em}

Never refuse a gift of betel leaf. It is an oath of hospitality. Refusing it is an insult that can end a bloodline.

\vspace{1em}

\noindent\textit{— The God-King's Astrologer, who has not seen the sun in thirty years}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice congee with dried shrimp—the rice is from the terraces, the shrimp is from the delta nets. The delta is where the river meets the sea. The sea is the God-King's mother. The God-King's mother is a nat. The nat is hungry.

\textbf{What does a child learn before they can walk?} To read the river. Ayokha children are taught the eight currents: Flood, Ebb, Eddy, Whirl, Still, and the ninth—the current that flows upstream. The upstream current is the monsoon's breath. The monsoon's breath is the God-King's voice. The God-King is silent.

\textbf{What is the first thing you notice about a stranger?} Their offering. Every traveler in Ayokha carries a gift for the nats—betel leaf, candle, a coin. A stranger whose offering is generous has been blessed. A stranger whose offering is small is testing the spirits. A stranger with no offering is a fool. The nats take fools first.

\textbf{What is the most common crime?} Nat-bribery—but bribery is not a crime, it is a negotiation. The nats set the tolls. The merchants pay the tolls. The nats spend the tolls on offerings. The offerings are burned. The ash falls on the fields. The fields are fertile. The circle is the economy.

\textbf{What does no one talk about but everyone knows?} The God-King is a puppet. The astrologer is the real ruler. The astrologer reads the stars. The stars say what the astrologer tells them to say. The astrologer has not read the stars in twenty years. He reads the nats' moods instead. The nats are fickle.

\textbf{Where do people go when they die?} To the river. The dead are cremated on the ghats. Their ashes are scattered on the water. The water carries them to the delta. The delta is the land of the nats. The nats decide whether the dead become ancestors or hungry ghosts. The nats are not generous.

\textbf{How do you apologize for a serious wrong?} You offer a nat—a spirit of the river, summoned by a medium. The nat hears your apology. If the nat accepts, the water is calm. If the nat refuses, the water rises. If the nat is silent, the water is still. Still water is the God-King's judgment. The God-King does not forgive.

\textbf{What is the most important festival?} The Welcoming of the Monsoon (summer). The God-King stands in the rain. The rain is the nats' blessing. The blessing is measured in inches. The inches determine the harvest. The harvest determines the taxes. The taxes determine the God-King's wealth. The God-King is not wealthy.

\textbf{What is the secret that could destroy a powerful person?} The God-King's daughter is not his daughter. She is the child of the astrologer and a nat—born from a bargain made during a failed monsoon. She can control the rain. She does not know. The astrologer knows. The astrologer is waiting for the right moment to tell her. The right moment is the next drought.
\end{almanac}

\newpage
\input{../generators/ayokah.tex}
\newpage

\chapter{Aelaerem of the Eastern Realms}
\label{chap:aelaerem-east}

\emph{The Hollow's Neighbors.}

\noindent
{\Large
\textit{My grandmother taught me that the Hollow is not evil. It is a forgetfulness. It forgets where the boundary lies, forgets that we are not food, forgets that it was once something else. So we remind it. With a thread, a bell, a bowl of salt. We are its memory.}

\vspace{0.5em}

I am a Thread-Binder. I tied a red thread around my wrist this morning. The Hollow walked past my door last night. It saw the thread. It kept walking. That is the bargain. I give it a boundary. It respects the boundary. Mostly. But last week it forgot the thread's colour. I had to tie two.

\vspace{0.5em}

The Aelaerem of the eastern realms have adapted their hedge-magic to thresholds of a different kind: bamboo groves that whisper in the dark, temple stairs that count your heartbeats, and the silence after a storm bell that has no clapper. Here, the Hollow wears the face of the uncanny — not hunger, but absence. It is the space between two bamboo stalks, the step that is always missing, the name you cannot quite remember.

They survive through ritual precision: a red thread tied at a crossroads with exactly eight knots, a bowl of rice left at dusk and never checked until dawn, a single bell rung once and then muffled with a cloth that has been washed in salt water. They whisper that the Hollow is the shadow of Kuva, cast when the first sun turned away, and that it wanders the world looking for the name it forgot.

Unlike western Aelaerem, who live in close-knit villages, the Eastern Aelaerem are often solitary or live in small family groups—yamabushi of the boundary, shrine hermits of the between. They are the ones who know where the veil is thinnest, and they are the ones paid to keep it closed. Their magic is humble: a knot to bind a promise, a thread to guide a lost ghost, a cup of sake to buy a night's silence.

\vspace{0.5em}

Never count the steps between your door and the crossroads. The Hollow counts with you. If you reach nine, it will be waiting at the ninth step. It will ask, “Where are you going?” Do not answer.

\vspace{1em}

\noindent\textit{— Hedge-Keeper Mila, who has never slept indoors}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
\textbf{What you notice first:} Red threads tied to every gate, every bridge, every well. The smell of burnt rice and old salt. The way the Aelaerem never whistle after dusk—the Hollow thinks you are calling it. And the way they never, ever count to nine.

\textbf{The one rule that kills newcomers:} Never count the steps between your door and the crossroads. The Hollow counts with you. If you reach nine, it will be waiting at the ninth step. It will ask, “Where are you going?” Do not answer.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Unspoken}

Power here is measured in red threads tied, offerings left, and the gaps between words. The Aelaerem have no leaders, no courts, no written laws. Each hedge-keeper knows her own stretch of boundary. She knows where the Hollow walks, when it is restless, and what gesture will turn it aside. Her authority is not given by any king — it is earned by surviving the uncanny. The Aelaerem are not feared or respected; they are tolerated, because without them the Hollow would seep into every threshold. They are the neighbours of the thing that lives between the notes of a song.

\subsection*{Starting Location}

A crossroads at dusk, where a red thread is tied around a wooden post. An old Aelaerem woman is pouring a cup of sake onto the ground. She looks up: ``The Hollow walked last night. I left salt. It took the salt. It left a single red thread tied around my wrist. It moved. Sleep with that knot, and it will know where you are. Do you want to know where it went? It went toward your camp.''

\subsection*{The Almanac of Ordinary Days (Eastern Aelaerem)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with pickled vegetables—the rice is from the paddies, the vegetables are salted in clay pots buried at the boundary. The salt is blessed by the hedge-keeper. The vegetables are always a little bitter. The bitterness is a reminder.

\textbf{What does a child learn before they can walk?} To tie the eight boundary knots. The ninth knot is never tied—it is left loose for the Hollow to find. If a child ties the ninth knot, the Hollow will visit that night and ask for a name. The child's name.

\textbf{What is the first thing you notice about a stranger?} Their shadow. A stranger whose shadow is too short has been touched by the Hollow. A stranger whose shadow moves independently is already lost. A stranger with no shadow is the Hollow itself, wearing a borrowed face. Do not invite them in.

\textbf{What is the most common crime?} Thread-theft—stealing a red thread from a boundary marker. The thread is not valuable, but cutting it breaks a promise. The Hollow becomes restless. The thief is marked: their shadow will shrink one inch each night until they return the thread.

\textbf{What does no one talk about but everyone knows?} The Hollow is not a single thing. It is a place—the space between heartbeats, the pause before a sneeze, the silence after a bell. The Aelaerem do not guard against a monster. They guard against the gaps.

\textbf{Where do people go when they die?} To the salt. The dead are cremated, and their ashes are mixed with salt and scattered along the boundary. The salt draws the Hollow's attention away from the living. The dead become the new boundary. They do not rest. They watch.

\textbf{How do you apologize for a serious wrong?} You leave a cup of sake and a lock of hair at the place where the wrong was done. The Hollow will witness. If the sake is gone by dawn and the hair is untouched, you are forgiven. If the hair is gone, the Hollow has taken your apology as a promise. You are now bound. Do not break that promise.

\textbf{What is the most important festival?} The Retting of the Threads (spring). All red threads on the boundary are taken down, washed in salt water, and re-tied. The village is silent for one night—no bells, no songs, no whispers. The Hollow listens. The Hollow remembers. If a thread is tied wrong, the Hollow will correct it. The correction is always a death.

\textbf{What is the secret that could destroy a powerful person?} The first red thread was not tied by an Aelaerem. It was tied by the Hollow itself, before it forgot what it was. The thread is still there, at the First Boundary. If it is untied, the Hollow will remember its true name. Then it will stop being hungry. It will become something else. No one knows what.
\end{almanac}

\newpage
\input{../generators/aelaerem_east.tex}
\newpage

\chapter{Aelinnel of the Inland Sea}
\label{chap:aelinnel}

\emph{The Bell-Tenders.}

\noindent
{\Large
\textit{The Hollow is not a place. It is a miscount. The ninth step that is never there, the ninth cup that is never poured, the ninth name that is never spoken. We do not fight the Hollow. We leave one empty chair, one unlit lamp, one uncounted stone. That is the price of sitting at the table.}

\vspace{0.5em}

I asked an Aelinnel bell-tender how she kept the tide-spirits from flooding her village. She tapped her temple. “I count them,” she said. “One, two, three, four, five, six, seven, eight. Then I stop. They know that if I ever count nine, I am not counting them.”

\vspace{0.5em}

The Aelinnel of the East do not hide from the spirits — they coexist. Small spirits inhabit every rock, tree, and wave. The Aelinnel have learned to negotiate with them using copper bells, precise counting, and occasional offerings of pickled plums. They are the engineers of shrines, the builders of floating bridges, and the keepers of the bell-lines that keep tide-spirits from straying. Some say Kuva, the sky, taught them the first counting song, and that the Hollow is simply the space where the count fails.

Unlike their western kin, who lean toward the Gift of the Mind, the Eastern Aelinnel practice a folk mathematics that blends logic and superstition. They believe every spirit has a number-name — a frequency that can be rung out of a bell, a sequence of steps that can be walked, a knot that can be tied. To offend a spirit is to miscount.

\vspace{0.5em}

Never count to nine aloud near a shrine. The spirits count with you. When you reach nine, they will ask: “What comes after nine?” The answer is always the Hollow.

\vspace{1em}

\noindent\textit{— Bell-Warden Tumble, who writes letters on copper wax}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
\textbf{What you notice first:} The sound of copper bells — not ringing in sequence, but chiming in overlapping waves. The way every bridge has a missing ninth plank. The smell of pickled plums and salt.

\textbf{The one rule that kills newcomers:} Never count to nine aloud near a shrine. The spirits count with you. When you reach nine, they will ask: “What comes after nine?” The answer is always the Hollow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Count}

Power here is measured in bells rung, tides predicted, and the precision of a threshold. The Aelinnel have no kings or councils; their authority rests on the bell-wardens who maintain the shrines and the tide-counters who read the waves. A village that loses its bell-warden is a village that the tide-spirits will flood. A bridge without its ninth plank missing is a bridge that the Hollow will cross. The Aelinnel do not worship the spirits; they transact with them. An offering is not a prayer — it is a payment. A bell is not a musical instrument — it is a contract.

\subsection*{Starting Location}

A floating shrine on the Inland Sea, where copper bells hang from every corner. A Bell-Warden rings a bell eight times, then stops: “The tide-spirit is restless. It wants a ninth bell. I will not give it. You — count the waves. Tell me how many break before the sun sets. If you miscount, the spirit will ask you for a story. Do not miscount.”

\subsection*{The Almanac of Ordinary Days (Aelinnel)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with pickled plums — the plums are fermented in salt brine and stored in copper-lidded barrels. The brine is never poured out; it is passed down through generations. The oldest batch is said to be nine generations old. No one has tasted it.

\textbf{What does a child learn before they can walk?} To count the eight safe numbers. Children are taught to count pebbles, footsteps, and bell-rings — but always to stop at eight. The ninth is never spoken. A child who says “nine” aloud is sent to sit by the missing plank of the bridge until they forget they ever said it.

\textbf{What is the first thing you notice about a stranger?} Their fingers. Aelinnel tide-counters often have only eight fingers — the ninth was a miscount they paid for. A stranger with nine fingers has never made a mistake. A stranger with seven has paid a heavy debt. A stranger with no fingers is either a spirit or someone who counted to nine and lost everything.

\textbf{What is the most common crime?} Bell-theft — but not for the copper. A stolen bell can be rung at the wrong time, creating a miscount. The thief is not punished by law; they are given to the tide-spirit whose count was broken. The spirit decides the price.

\textbf{What does no one talk about but everyone knows?} The First Bell was not cast by an Aelinnel. It was found — already ringing, already counting. No one knows who made it. No one knows why it rang. The shards are scattered across the Inland Sea. When the ninth shard is found, the First Bell will ring again. The Hollow will hear it.

\textbf{Where do people go when they die?} To the bell-line. The dead are not buried. Their ashes are mixed with copper dust and cast into new bells. The bells are hung on the bell-lines. When the tide-spirits ring them, the dead speak. Mostly they say: “Count carefully.”

\textbf{How do you apologize for a serious wrong?} You offer a count — a public recitation of the wrong, spoken in eights. “I took the offering. I broke the bell. I counted to nine.” Then you wait. If a spirit rings a bell eight times, you are forgiven. If it rings nine times, you are cursed. If no bell rings, the Hollow will visit you at midnight.

\textbf{What is the most important festival?} The Retuning of the Bells (autumn). Every bell on the Inland Sea is struck once, in sequence, from the highest tide to the lowest. The sound travels for miles. The spirits listen. If a bell is out of tune, the spirits will correct it by morning. The correction is always a death — but the bell will ring true for another year.

\textbf{What is the secret that could destroy a powerful person?} The ninth bell was never meant to be cast. It was a mistake — a foundry worker poured one bell too many. They tried to hide it, but it rang once on its own. The tone was the Hollow's name. The bell still exists. It is buried under a shrine on a small island. No one knows which island. The map was miscounted.
\end{almanac}

\newpage
\input{../generators/aelinnel_inland_sea.tex}
\newpage

\chapter{Lethai-al of Nihori}
\label{chap:lethai-al}

\emph{The Storm-Touched Wanderers.}

\noindent
{\Large
\textit{We do not own the land. We walk it. The road is our shrine. The storm is our law. The sea does not forgive, but it does forget. If you can wait long enough, the waves will erase your debt. But we do not wait. We walk. From shrine to shrine, from storm to storm, we carry the names of the drowned. The road is a prayer. Every step is a syllable.}

\vspace{0.5em}

My grandmother drowned in the ninth wave. Her name became a tide-charm. I keep it in a pouch of oiled leather. When the storm grows angry, I whisper her name. The sea remembers her. It calms for a moment — long enough to breathe. We are wanderers. The road is our home. The dead travel with us.

\vspace{0.5em}

The Lethai-al of Nihori once guarded a drowned kingdom — an archipelago that sank when a forbidden name was spoken. Their blood still carries the salt of that ancient catastrophe. Now they are wanderers in the oldest sense: masterless, landless, bound to no lord but fiercely bound to each other and to the storm spirits that swirl above the waves. They walk from shrine to shrine, from port to port, following the processional roads that circle the island.

They do not farm. They do not build. They carry everything they own on their backs or in wooden carts. Their duels are not to the death — they are to the first draw of blood, after which the loser must join the procession and carry the winner's burden for a season. Their courts are crossroads. Their temples are the road itself. The Eastern Ninth lives here: the Hollow does not hide under hills — it breathes in the salt foam. The ninth wave is never the largest; it is the one that remembers your name.

\vspace{0.5em}

Never walk under a torii gate without making an offering. The gate is a threshold. The road on the other side is not the same road you left.

\vspace{1em}

\noindent\textit{— Procession Elder Kenji, who has walked the circuit 99 times}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
\textbf{What you notice first:} The smell of salt and incense. The way the Lethai-al never sit with their backs to the sea. The constant, low murmur of waves being predicted — not by instruments, but by the tilt of a head, the twitch of a finger. The worn wooden plaques hung from their belts, each carved with a dead ancestor's name.

\textbf{The one rule that kills newcomers:} Never walk under a torii gate without making an offering. The gate is a threshold. The road on the other side is not the same road you left.
\end{tcolorbox}

\subsection*{Regional Mood: The Processional Road}

Power here is measured in shrines visited, storms survived, and the number of dead ancestors carried as tide-charms. The Lethai-al have no king, no courts, no written law. Their authority is personal and processional: a Procession Elder's word is law because she has walked the circuit more times than anyone else. Their guilds are not guilds — they are processional bands, each with its own rhythm, its own counting song, its own patron shrine. A band that walks too fast offends the spirits; one that walks too slow loses the tide. The Lethai-al do not believe in ownership of land; they believe in passage rights, storm oaths, and the memory of the drowned. Every crossroads is a court. Every storm is a messenger. Every sunrise is a departure.

\subsection*{Starting Location}

A procession trail winding along the sea cliffs, marked by weathered torii gates and small stone shrines. A group of Lethai-al walk in single file, chanting a counting song. One carries a palanquin with a covered object — a tide-charm shrine. The leader stops: “You stand on the processional road. The drowned are watching from the waves. Do you walk with us? If so, you must carry nothing. The road provides. The road takes away.”

\subsection*{The Almanac of Ordinary Days (Lethai-al)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Dried fish and rice balls — the fish is smoked over driftwood fires, the rice is pressed into shapes that fit in a pouch. The food is always cold. The Lethai-al do not stop to cook. The road does not wait.

\textbf{What does a child learn before they can walk?} To count the nine waves. Children are taught to watch the sea and count: one, two, three, four, five, six, seven, eight — then stop. The ninth wave is the spirit's wave. It counts you. If you count it, you owe it.

\textbf{What is the first thing you notice about a stranger?} Their feet. A Lethai-al who has walked the circuit for many years has calloused soles cracked like dry earth. A stranger with soft feet has not walked far. A stranger with no feet is a ghost who lost the road.

\textbf{What is the most common crime?} Charm-theft — stealing a tide-charm from a sleeping traveller. The thief gains the dead ancestor's favour for one night. The ancestor wakes. The ancestor is angry. The punishment is to carry the victim's burden for the rest of the circuit.

\textbf{What does no one talk about but everyone knows?} The drowned kingdom did not sink by accident. The forbidden name was spoken by a Lethai-al — a woman who lost her child to the sea. She wanted the sea to remember. It remembered too well. Her name is not spoken. Her tide-charm hangs in the Sunken Gate, still wet.

\textbf{Where do people go when they die?} Into the tide-charms. The dead are not buried. Their names are carved onto shells and placed in oiled leather pouches. The living carry them. The dead walk with the living. When the pouch is lost, the dead become the Hollow.

\textbf{How do you apologize for a serious wrong?} You walk a crossroads circuit — nine times around a crossroads shrine, counting your steps in eights. On the ninth circuit, you must speak the wrong aloud. If the wind carries your voice away, you are forgiven. If the wind throws it back, you are not. If the wind is silent, the Hollow is listening.

\textbf{What is the most important festival?} The Procession of the Ninth Wave (autumn). All Lethai-al bands walk to the shore at low tide. When the ninth wave crashes, they wade into the foam and whisper the names of those who drowned that year. The sea answers with a single wave. The wave is always the same height. The height is the number of names.

\textbf{What is the secret that could destroy a powerful person?} The Ninth Wave Herald is not a child. She is the first drowned — the woman who spoke the forbidden name. She did not die. She became the ninth wave. She returns every year to point at the horizon. She is pointing at the day the sea will rise again. No one knows which day.
\end{almanac}

\newpage
\input{../generators/lethai_al_wanderers.tex}
\newpage

\chapter{Lethai-ar of Ayokha}
\label{chap:lethai-ar-ayokha}

\emph{The River-Bound Oath-Keepers.}

\noindent
{\Large
\textit{A witness cord is not a leash. It is a mirror. If you speak truly, you see yourself. If you lie, you see what you could have been. I have cut two cords in my life. I still dream of the faces in the silk. The river remembers every word spoken on its banks. We are merely the scribes.}

\vspace{0.5em}

My lamp has burned for nine years without being refilled. The flame is green. It flickers when a merchant lies about the weight of his silk. The river does not need oil. It needs witnesses. I sit on this wharf and watch. That is my offering.

\vspace{0.5em}

The Lethai-ar of Ayokha are the living bridges between mortals and the river spirits that launder the world's promises. Their ancestors were river-folk that ascended to the realm of the fair folk, and they now walk the middle path, a step between bodies of water and the world of men. They keep lanterns that never go out, record every contract spoken on the monsoon wharves, and bind their vows with silken cords dyed in river mud. Unlike their northern kin who swore to distant deities, the Lethai-ar serve \textbf{Benzaiten} (the river of words) or \textbf{Inari} (the fox's ledger). Their vows are physical cords, braided from their own hair and dyed with river mud. When a vow is fulfilled, the cord is burned. When it is broken, the cord tightens around the oath-breaker's wrist until blood is drawn. The Ayokhan taboo against the ninth cup is enforced by the Lethai-ar: the ninth cup is for the river spirits, and pouring it is an invitation to the Hollow.

\vspace{0.5em}

Never pour the ninth cup of tea or sake in a Lethai-ar's presence. The ninth cup is for the witness. Pour it, and you have invited the river to judge you. The river is not merciful.

\vspace{1em}

\noindent\textit{— Vigil Lady Suyin, who has cut nine witness cords — the ninth was her own}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
\textbf{What you notice first:} The smell of river water and oiled silk. The way every Lethai-ar carries a small oil lamp, even at noon. The constant murmur of contracts being witnessed — a merchant's promise, a captain's oath, a lover's vow.

\textbf{The one rule that kills newcomers:} Never pour the ninth cup of tea or sake in a Lethai-ar's presence. The ninth cup is for the witness. Pour it, and you have invited the river to judge you. The river is not merciful.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Witness Cord}

Power here is measured in lanterns kept lit, cords tied, and the number of oaths witnessed. The Lethai-ar have no army, no treasury, no territory. They have the river, and the river remembers everything. Every contract in Ayokha requires a Lethai-ar witness to be legally binding. The God-King's astrologers may set the calendar, but the Lethai-ar hold the ledger. Their neutrality is absolute: they will witness a vow between enemies as readily as between allies. The price of a witness cord is not silver — it is a promise to keep the river clean, to light a lamp at dusk, to speak no word that cannot be unsaid.

\subsection*{Starting Location}

A monsoon wharf on the great river, where silk barges are being loaded. A Vigil in grey robes holds a lantern of green flame: “The merchant and the captain have made a vow. The captain says the silk is eight bales. The merchant says it is nine. The witness cord is tied. It will tighten around the liar's wrist. I do not need to know who is lying. The cord will tell me. But you — you are not bound. Watch. Learn. And do not pour the ninth cup.”

\subsection*{The Almanac of Ordinary Days (Lethai-ar of Ayokha)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with river fish — the fish are caught in the sacred currents, and the bones are returned to the water as an offering. The porridge is salted with evaporated river brine. It tastes of memory.

\textbf{What does a child learn before they can walk?} To tie the eight witness knots. Each knot represents a clause of a fair contract. The ninth knot is never tied — it is left for the river to judge. A child who ties the ninth knot is taken to the step-pyramid and made to watch the green flame until they can recite every oath their family has ever sworn.

\textbf{What is the first thing you notice about a stranger?} Their lamp. A Lethai-ar’s lamp is a public record of their witness. A stranger whose lamp burns green has kept every oath. A stranger whose lamp flickers has witnessed a broken vow. A stranger whose lamp is dark has been cut from the Vigil. Do not speak to them. The river may be listening.

\textbf{What is the most common crime?} Cord-forging — creating a false witness cord that mimics the dye of a respected Vigil. The forger is not executed. Their hands are bound with a cord that tightens each time they speak. If they never lie, the cord will eventually loosen. If they lie once, it will cut.

\textbf{What does no one talk about but everyone knows?} The First Cord was not woven by a Lethai-ar. It was found — already knotted, already witnessing — floating down the river on a broken barge. No one knows who tied it. No one knows what oath it witnessed. The cord is kept in a lead-lined box at the bottom of the Well of Broken Oaths. The box is warm. The cord is still waiting.

\textbf{Where do people go when they die?} To the green flame. The dead are cremated, and their ashes are mixed with lamp oil. The oil is poured into the Vigil lanterns. The dead become the light. The light witnesses. The light does not lie.

\textbf{How do you apologize for a serious wrong?} You offer a new cord — woven from your own hair, dyed with your own blood. The cord is tied between you and the wronged party. If the cord tightens, you are not forgiven. If it loosens, you are. If it breaks, the river will decide your fate. The river is patient.

\textbf{What is the most important festival?} The Reckoning of the Cords (spring). All witness cords that have been cut or broken are taken from the Well of Broken Oaths and burned in a barge on the river. The smoke rises. The river spirits watch. If a cord does not burn, the oath it witnessed is still unpaid. The debtor has one more year. If the same cord fails to burn twice, the Hollow comes.

\textbf{What is the secret that could destroy a powerful person?} The God-King’s own witness cord is a forgery. He swore a vow to protect the Lethai-ar nine years ago. His cord was cut the next day. The Vigil knows. They have not told anyone because the cord that cut his was their own. The ninth cord that Lady Suyin cut was the king’s. She has been waiting for him to break his oath ever since. He will. They always do.
\end{almanac}

\newpage
\input{../generators/lethai_ar_river.tex}
\newpage

\chapter{Lethai-thora of Sihai}
\label{chap:lethai-thora}

\emph{The Silent Ledgers.}

\noindent
{\Large
\textit{We do not burn the dangerous scrolls. We bury them in plain sight, on shelves where no one thinks to look. The Emperor's Censor once stood in our archive and asked for a tax ledger. We gave him one. He did not notice the shelf behind him held the true history of his own grandfather's death. Silence is not emptiness. It is a door that opens only inward.}

\vspace{0.5em}

My grandmother tied this scroll with a silk cord. The cord has nine knots. The ninth knot is a secret she never told me. When I die, I will tie a tenth. Someone will untie it in a century. They will learn why the emperor's concubine wept on the night of the lantern festival. Some secrets are too heavy for one lifetime.

\vspace{0.5em}

The Lethai-thora were once courtiers of a celestial fey realm that spanned the Valewood and the eastern hills. When the old fey courts fell — when the Alder King withdrew and the Hazel Queen sealed her gates — many Lethai-thora fled east. They crossed the mountains with nothing but armfuls of scrolls and the memory of a civilisation that had forgotten how to die.

In the fertile river valleys of what would become Sihai, they found scattered agrarian tribes fighting over water and silence. The Lethai-thora did not conquer. They taught. They taught the tribes to write contracts instead of drawing blood, to measure grain instead of stealing it, to record debts instead of nursing grudges. Over generations, the tribes became a kingdom, the kingdom became an empire, and the empire became Sihai. The Lethai-thora became its memory.

Now they dwell in hidden pavilions, preserving “forgotten ledgers” that record what the Mandate of Heaven prefers to forget. Their knowledge is a burden: each secret they preserve slowly erases a piece of themselves. A Lethai-thora who guards a forbidden truth may find that she can no longer recall the colour of her mother's dress — a small price, she says, for keeping the ledger whole. They are no longer fey. They are no longer fully human. They are something in between: scribes of a story that has no ending, archivists of a past that never quite happened.

\vspace{0.5em}

Never touch a scroll without announcing your lineage. Even a false lineage is better than silence. The scrolls remember who opened them, and they will whisper your name to the Hollow.

\vspace{1em}

\noindent\textit{— Keeper Lian, who has not spoken aloud in twenty years — she writes on water}
}

\vspace{1em}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere}]
\textbf{What you notice first:} The smell of old parchment and incense. The way scrolls are stacked with their spines facing the wall — so the words cannot see who reads them. The silence when a secret is spoken; even the insects stop chirping.

\textbf{The one rule that kills newcomers:} Never touch a scroll without announcing your lineage. Even a false lineage is better than silence. The scrolls remember who opened them, and they will whisper your name to the Hollow.
\end{tcolorbox}

\subsection*{Regional Mood: The Weight of the Forgotten}

Power here is measured in scrolls sealed, secrets kept, and the length of a silence. The Lethai-thora have no king, no army, no treasury. They have archives, and the archives are their fortress. The Emperor's Censorate fears them because they cannot audit a secret that has no ink. The Mandate of Heaven is written in public; the Lethai-thora keep the unwritten ledger — the one that records why dynasties fall. They are not rebels; they are witnesses. And a witness who never speaks is the most dangerous kind.

\subsection*{Starting Location}

A hidden pavilion behind a waterfall, where scrolls are stacked to the ceiling and the air is dry as bone. A Keeper of the Seventh Gutter looks up from a brass lamp: “The Emperor's Quiet Chancellor has sent a messenger. He wants a scroll that does not exist. He wants the truth about his mother's death. The truth is written in water. It will vanish if we speak it. But you — you are not bound by our silence. Will you carry a memory for us?”

\subsection*{The Almanac of Ordinary Days (Lethai-thora)}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with dried fish — the fish are smoked over fires fed with old scrolls. The smoke smells of forgotten contracts. The porridge is always lukewarm. The archives are never truly warm.

\textbf{What does a child learn before they can walk?} To read without moving their lips. A Lethai-thora child who whispers a secret aloud is sent to sit in the Chamber of the Ninth Scroll until they learn that silence is a shield. Some children never come back. They become scribes of the empty shelves.

\textbf{What is the first thing you notice about a stranger?} Their ink‑stained fingers. A Lethai-thora has ink under every fingernail. A stranger whose fingers are clean has never copied a forbidden text. A stranger whose fingers are stained black has copied too many. A stranger whose fingers are stained red has copied a death warrant — their own.

\textbf{What is the most common crime?} Context key theft — stealing the carved stone that unlocks a scroll's meaning. The thief is not punished by law. They are given a blank scroll and told to write their own confession. The scroll will fill itself if the thief lies. The ink is their own blood.

\textbf{What does no one talk about but everyone knows?} The Ninth Scroll is not a scroll. It is a memory so dangerous that nine different keepers each sealed a fragment of it in their own minds. When the ninth keeper dies, the memory will be lost forever — or will escape.

\textbf{Where do people go when they die?} To the Well of Forgotten Names. The dead are not buried. Their bodies are placed in the well, and the well swallows them. The water turns black. The black is the ink of their unwritten secrets. The well is bottomless. It is never full.

\textbf{How do you apologize for a serious wrong?} You offer a silence — a period of total stillness, no breath, no heartbeat, no word. The silence is timed by a water timer. If you hold the silence for nine drops, you are forgiven. If you break it, you must speak the wrong aloud in the Chamber of the Ninth Scroll. The scroll will remember. The scroll does not forgive.

\textbf{What is the most important festival?} The Unrolling of the Ledgers (winter). Once a year, the keepers unroll a single forbidden scroll and read it aloud to each other — but they read it in a whisper, and only the walls hear. The walls remember. The walls do not judge. The festival ends with all nine keepers drinking ink from a silver cup. The ink is their own memory. The next year, they will have forgotten.

\textbf{What is the secret that could destroy a powerful person?} The emperor's true lineage is recorded on a scroll that no living keeper has opened. The scroll is sealed with a cord tied by the first keeper. The cord has nine knots. The ninth knot is a lie — the emperor's grandfather was not of the blood. The empire was founded by a concubine's son. The scroll is warm. It is kept in the First Pavilion. The First Pavilion is invisible. Only those who have forgotten nine names can see it.
\end{almanac}

\newpage
\input{../generators/lethai_thora_east.tex}
\newpage

\part{The Spaces Between}
% After \part{The Spaces Between}
\begin{quote}
  \small
  \textit{``The Ways Between do not measure miles. They measure choices. A wrong turn might lead to yesterday. A right turn might lead to a version of yourself who made a different decision – and is still alive to regret it. The mist does not hide landmarks. It hides destinations.''}
  \noindent — Dusana of the Raven Road, at a crossroads that did not exist on any map
  \end{quote}
% ------------------------------
% CHAPTER: UBRAL
% ------------------------------
\chapter{Ubral}
\emph{The Stone Between Spears.}

\noindent
{\Large
\textit{The stone here remembers every oath sworn upon it and every blood spilled in its shadow. To lead is to understand that the land itself is the true sovereign, and we are merely its stewards.}

\vspace{0.5em}

I am a Hearth-Aunt. Hospitality's a sacred thing in these hills—break bread at my table and you're family till the sun sets twice. But cross that trust and the very stones will remember your name.

\vspace{0.5em}

Ubral is a land of upland mists, stone cairns, and clan oaths held tighter than iron. Caught between the Aelerian mountains and the lowland courts of Viterra, its hills are scarred by old raids and crowned by watch-fires that speak faster than any rider. Here, law is written not in charters but in cattle, wergild, and songs; a guest's word can buy more than a sword, but one broken vow can spark a feud lasting generations. But here's the secret the Lady of Tor doesn't want you to know: the cairns are not just markers. They are witnesses. Every stone in every cairn has a name. If you touch the right stone, you can hear the dead voting on the disputes of the living. The dead have long memories. They vote with the wind.

\vspace{0.5em}

Viterra has tried to conquer Ubral a dozen times; each attempt ended with supply lines cut by reivers, passes sealed by Aeler outcast engineers, and lowland knights lost in mist. The Utaran Imperium itself broke its leg on these hills and limped away. Now the clans watch the lowlands with patient eyes, and the exiled Aeler of Khaz-Vurim weigh every traveler's honesty on iron scales. A guest here is kin—until they prove otherwise. And kin here will die for you—or kill you over a stolen cow.

\vspace{0.5em}

Never refuse a cup from a hearth-aunt. Hospitality is a contract sealed in salt and bread. Refusal is an insult that blood must answer.

\vspace{1em}

\noindent\textit{— Hearth-Aunt Brigid, who has never drawn a blade and has never needed to}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Oatcakes and skyr—the oats are grown in highland clearings, the skyr is cultured in stone crocks buried in the peat. The peat is ancient. The peat remembers the old kings. The old kings are buried beneath it.

\textbf{What does a child learn before they can walk?} To count cattle. Ubral children are taught the eight breeds of the hills: Black, Red, White, Brindled, Dun, Grey, Spotted, and the ninth—the breed that has no colour because it belongs to the cairn-folk. The cairn-folk are the dead. The dead do not pay taxes.

\textbf{What is the first thing you notice about a stranger?} Their guest-token. Every Ubral traveller carries a token—a stone, a bone, a braid of hair—given by their last host. A stranger whose token is warm has been recently welcomed. A stranger whose token is cold has been travelling long. A stranger with no token is a spy. Spies are given a token—a rope. The rope is for their neck.

\textbf{What is the most common crime?} Cattle raiding—but raiding is not a crime, it is a sport. The cattle are the wealth. The wealth is the clan's. A stolen cow is a stolen child. The child must be returned. The return is a negotiation. The negotiation is a feud. The feud is a song.

\textbf{What does no one talk about but everyone knows?} The Lady of Tor is not the ruler. The cairns rule. The cairns vote at each full moon. Their votes are carried by the wind. The wind speaks to the hearth-aunts. The hearth-aunts speak to the clan. The clan obeys. The Lady is the voice. The cairns are the vote.

\textbf{Where do people go when they die?} To the cairns. The dead are not buried. Their bones are added to the cairns. The cairns grow taller each year. The tallest cairn is the oldest. The oldest cairn has no name. The oldest cairn remembers the first clan.

\textbf{How do you apologize for a serious wrong?} You offer a wergild—a payment in cattle, calculated by the wronged party's estimate of their loss. The cattle are driven to the cairn. The cairn is the court. If the cattle graze peacefully, you are forgiven. If they scatter, you are not. If they refuse to enter the cairn's shadow, your debt is blood.

\textbf{What is the most important festival?} The Gathering of the Clans (autumn). The clans meet at the Great Cairn. The cattle are counted. The songs are sung. The feuds are settled. The dead are named. The wind carries the names to the hills. The hills remember.

\textbf{What is the secret that could destroy a powerful person?} The Great Cairn is not a cairn. It is a prison—the tomb of a Viterran king who tried to conquer Ubral and failed. He was buried alive, with his crown and his sword. His ghost votes with the cairns. He votes for war. The cairns have not yet agreed.
\end{almanac}

\newpage
\input{../generators/ubral.tex}
\newpage

% ------------------------------
% CHAPTER: SILKSTRAND
% ------------------------------
\chapter{Silkstrand}
\emph{City of Bridges \& Dyewater.}

\noindent
{\Large
\textit{In this city, every bridge is a contract and every current carries coin. The bargains struck in our shadows never quite fade from memory, and the waters remember every handshake.}

\vspace{0.5em}

I am a Bobbin-Runner. I can cross this city faster than any grown-up, and I know which planks creak and which merchants lie. The bridges remember every step, and the canals carry more secrets than fish.

\vspace{0.5em}

Silkstrand is a city strung across canals and arches, a place where every bridge is a contract and every current carries rumor. The silk looms never sleep, the Exchange never closes, and the Archivolt's notaries seal fates with a drop of wax. Merchants rise and fall with the tide of the market, bravos carve reputations on the planks of Three-Queens Bridge, and the city's underworld runs on whispers and borrowed keys. But here's the secret the Matron doesn't want you to know: the canals are not just water. They are memory. Every bribe, every broken promise, every cargo that was never declared—the canals remember. And sometimes, at low tide, the water shows you the faces of the merchants who drowned rather than pay their debts.

\vspace{0.5em}

The Matron rules from her palazzo stairs with velvet and iron, but the city itself belongs to silk, silver, and shadow. The Exchange Volatility Timer ticks with every major cargo seizure, every trade route cut, every merchant default. When it fills, a panic hits—the Exchange closes, and all outstanding contracts are voided unless witnessed by the Matron herself. The Matron has voided nine contracts this year. The ninth was her own.

\vspace{0.5em}

Never name a rival while crossing a bridge at night. The water remembers, and the curse you send may come back wrapped around your own throat.

\vspace{1em}

\noindent\textit{— Bobbin-Runner, who ties a knot in her apron string for every secret she carries}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with dried fish—the rice is from the Oshiiran barges, the fish is from the Zakov fleets. Both are smuggled. Both are delicious. Both are paid for in favours.

\textbf{What does a child learn before they can walk?} To count bridges. Silkstrand children are taught the eight bridges of the city: Merchant's, Weaver's, Dyer's, Cooper's, Pilot's, Beggar's, Saint's, and the ninth—the bridge that is only there at low tide. The ninth bridge leads to the Sunken Quarter. The Sunken Quarter is where debts go to drown.

\textbf{What is the first thing you notice about a stranger?} Their ink. Every Silkstrand merchant has a seal-ink stain on their right thumb. A stranger whose ink is fresh is new to the trade. A stranger whose ink is faded has been retired. A stranger with no ink is not a merchant. Not a merchant is not a person.

\textbf{What is the most common crime?} Contract fraud—but fraud is not a crime, it is a negotiation. The contracts are written in disappearing ink. The ink's formula is a secret. The secret is held by the Archivolt notaries. The notaries sell the antidote. The antidote is the fee.

\textbf{What does no one talk about but everyone knows?} The Matron is not the ruler. The canals rule. The canals set the prices. The canals decide who lives and who drowns. The canals are not water. The canals are the city's memory. The memory is the Matron's. The Matron is the canals' voice.

\textbf{Where do people go when they die?} To the canals. The dead are not buried. Their bodies are wrapped in silk and sunk in the deep channels. The channels are the city's ledger. The ledger is never balanced.

\textbf{How do you apologize for a serious wrong?} You offer a bridge—a toll paid to the Bridge-Lords. The toll is calculated by the wronged party's estimate of their loss. The toll is paid in silver. The silver is melted and poured into the canal. The canal remembers the weight. The weight is your apology.

\textbf{What is the most important festival?} The Dyeing of the Waters (spring). The dyers empty their vats into the canals. The water turns red, blue, green, gold. The colours are the year's fashions. The fashions are set by the Matron. The Matron's favourite colour is the one that sells best. The best colour is the one she dyed her own hair.

\textbf{What is the secret that could destroy a powerful person?} The canals are not natural. They were dug by a Sidhi engineer who was paid in promises. The promises were broken. The engineer cursed the city. The curse is that every ninth drop of water is a tear. The tears belong to the engineer. She is still crying. She has been crying for nine centuries.
\end{almanac}

\newpage
\input{../generators/silkstrand.tex}
\newpage

\chapter{Midh-Akhaz}
\emph{The Violet Gate.}

\noindent
{\Large
\textit{Coin talks. Grass walks. Blood settles the rest.}

\vspace{0.5em}

I am a Horse-Braid Apprentice. I tie a new knot for every paying passenger. My hair is a map of debts — and grudges. The wind carries names, not ledgers, but here, in the shadow of the Oasis Compact, both trade and oaths change hands. Do not mistake our patience for weakness. The grass remembers every hoof that crossed it.

\vspace{0.5em}

Midh Ahkaz is not a city. It is a wound that learned to trade. A day's ride north of the Ecktorian border, it squats on the edge of the Violet Steppe like a patient wolf. No walls. No gates. Just rings of felt tents, latrine trenches, and the smell of horse sweat, cedar smoke, and the faint copper tang of blood money.

\vspace{0.5em}

The grass grows tall enough to hide a body. The locals cut it back at dawn — but never at dusk, because the things that crawl through it at night have sharp teeth and no patience for property lines. Ecktorian emissaries call this place ``the Brink.'' Caravan guards call it home. Thieves call it paradise.

\vspace{0.5em}

Every morning, the Coin-Weigh Tribunal opens its scales under a canvas dome. Every night, the Dust Market auctions goods that fell off a wagon — or off a corpse. The Khatun's word is law until someone stronger speaks. And in the Silver Stall Stable, a horse thief can buy a fresh mount with a story and a knife.

\vspace{0.5em}

Never refuse the guest‑cup. The first pour is courtesy. The second is negotiation. The third is a blood‑price unspoken. And the fourth — the fourth means you're already dead, you just haven't stopped breathing.

\vspace{1em}

\noindent\textit{— Horse-Braid Apprentice, who ties a knot for every secret she carries}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Curd cheese and flatbread — the cheese is from the Ykrul herds, the bread is baked in the embers of a dung fire. It tastes of smoke and the open sky. The poor dip it in salt; the rich dip it in honey stolen from a border hive.

\textbf{What does a child learn before they can walk?} To count horses. Midh Ahkaz children are taught the eight gaits of a healthy mount: walk, trot, canter, gallop, pace, amble, rack, and the ninth — the death‑sprint that only a stolen horse knows. The ninth gait is never taught; it is discovered when the rider's life depends on it.

\textbf{What is the first thing you notice about a stranger?} Their braids. Ykrul warriors braid their hair with clan colours; Vilikari traders wear oiled braids with silver beads. A stranger whose braids are loose has just fought. A stranger whose braids are cut has lost a blood‑price. A stranger with no braids is either a slave or a ghost. Neither stays long.

\textbf{What is the most common crime?} Grass‑right forgery — but forgery is not a crime, it is a negotiation. The Oasis Compact issues branded sticks for grazing rights. A clever carver can duplicate the brand. The penalty is exile into the tall grass at dusk. The grass does not have a court of appeals.

\textbf{What does no one talk about but everyone knows?} The Khatun is not the only ruler. The Silver Stall Stable master knows where every stolen horse sleeps. The Dust Market auctioneer knows the price of every sin. And the Iron Well keeper knows the name of everyone who will die before the next grass moon. The three of them meet once a season. No one records what they say. The grass records it.

\textbf{Where do people go when they die?} To the grass. The dead are not buried. Their bodies are carried beyond the caravan rings and left for the wolves and the wind. The grass grows tall and green where a body falls. A patch of deep violet is a sign that the dead were not properly mourned. The locals avoid those patches.

\textbf{How do you apologize for a serious wrong?} You offer a blood‑price in silver, weighed on the Coin‑Weigh Tribunal's scales. If the scale tips in your favour, you are forgiven. If it tips against you, you owe a horse. If the scale balances, you owe a story — and the story must be true, and the true story must make the wronged party weep. If it does not, you owe a finger.

\textbf{What is the most important festival?} The Grass Moon (autumn equinox). The violet seed‑heads glow under moonlight. The tribes ride in a great circle, singing the names of the dead. At midnight, the Khatun pours a cup of water onto the earth. If the water sinks, the winter will be mild. If it pools, the winter will be hungry. If it steams, the Hollow is walking. The festival ends at dawn. No one sleeps.

\textbf{What is the secret that could destroy a powerful person?} The Oasis Compact's original seal was not lost. It was stolen by a Vilikari factor who wanted to control water rights. The factor is still alive, living in the Foreign Quarter under a false name. The seal is hidden in a hollow horse skull buried under the Grass Ring. The skull belonged to the Khatun's favourite stallion. The stallion did not die of old age.
\end{almanac}

\newpage
\input{../generators/midh-akhaz.tex}
\newpage

% ------------------------------
% CHAPTER: THE WAYS BETWEEN
% ------------------------------
\chapter{The Ways Between}
\emph{Spiritways \& Veilways.}

\noindent
{\Large
\textit{You are reading this sentence twice. The first time, you understood it. The second time, it understood you.}

\vspace{0.5em}

I am a Lost Pilgrim. Or I was. Or I will be. The paths here don't care about tenses. They care about which version of you arrives at the end. I've been walking for—no. That's a lie. The walking has been walking me. I forget which of us started first.

\vspace{0.5em}

The Ways Between do not measure miles. They measure choices. A wrong turn might lead to yesterday. A right turn might lead to a version of yourself who made a different decision and is still alive to regret it. The mist does not hide landmarks. It hides destinations. And sometimes, when you walk long enough, you realise the destination was never the point. The point was the moment you turned left instead of right. The point was the breath you held at the crossroads. The point was the name you did not speak.

\vspace{0.5em}

Here is the secret the Wayfinder doesn't want you to know: there is no Wayfinder. There is only the road, and the road is a liar. It tells you it wants you to find your way. What it actually wants is to watch you choose. Every reflection in every pool is a version of you that chose differently. They are all still walking. They are all still watching. And they are all still waiting for you to make the choice that finally lets them stop.

\vspace{0.5em}

I met myself on a bridge once. They were coming from the opposite direction. We stopped. We looked at each other. We said nothing. We both knew that if one of us spoke, the other would have to listen. And listening is a kind of debt.

\vspace{0.5em}

Never answer a question asked by a reflection. The answer will bind you to the path you did not choose. But here is the second secret: not answering is also a choice. And the road is very, very patient.

\vspace{1em}

\noindent\textit{— The Wayfinder, who exists only in the moment a choice is made, and who is not the same person at the end of this sentence as at the beginning}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Nothing. There is no breakfast in the Ways Between. There is no morning. There is no hunger. There is only the walking. The walking is sustenance.

\textbf{What does a child learn before they can walk?} To forget. Children born in the Ways Between do not learn. They remember. They remember the path they took to be born. The path has eight turns. The ninth turn is the birth. The birth is a choice. The child does not know they chose.

\textbf{What is the first thing you notice about a stranger?} Their direction. A stranger walking toward you may be coming from your future. A stranger walking away may be going to your past. A stranger walking beside you may be yourself from a timeline where you did not turn left. Do not ask them which. They will not answer.

\textbf{What is the most common crime?} Stealing—but stealing is not a crime, it is a paradox. A thief in the Ways Between steals a memory from a version of you that no longer exists. The memory never belonged to you. You cannot miss it. You will miss it anyway.

\textbf{What does no one talk about but everyone knows?} The Ways Between are not between anything. They are the only thing that exists. The solid ground you remember—the cities, the roads, the hearths—is the dream. The Ways Between are the waking. The waking is the walking. The walking never ends.

\textbf{Where do people go when they die?} Deeper. The Ways Between have no surface, no bottom, no end. The dead walk deeper. The deeper they walk, the less they remember. When they remember nothing, they become the path. The path is the walking. The walking is the dead.

\textbf{How do you apologize for a serious wrong?} You cannot. Apologies require memory. Memory requires a self. The Ways Between erase both. The only apology is to stop walking. Stopping is not an option.

\textbf{What is the most important festival?} There are no festivals. There is only the moment between heartbeats when you realise you have been walking for nine years and you do not remember why. That moment is the festival. Celebrate it by taking the next step.

\textbf{What is the secret that could destroy a powerful person?} The Wayfinder is not a person. It is the road itself, dreaming of a traveller who will finally reach the end. The end does not exist. The road knows this. The road dreams anyway.
\end{almanac}

\newpage
\input{../generators/ways_between.tex}
\newpage

% ------------------------------
% CHAPTER: THE WILDS
% ------------------------------
\chapter{The Wilds}
\emph{Roads, Ruins, and Weather.}

\noindent
{\Large
\textit{Between towns and treaties lies the patience of the land. Old roads remember armies; cairns remember names no book keeps. In the Wilds, law thins to trail-marks and favors, and the weather keeps its own counsel. The Edgewalker calls it "unclaimed inventory." The land calls it patience. The land has more patience than the Edgewalker.}

\vspace{0.5em}

I am a ranger now. Stonehaven trained me as a scholar—mapping ruins, copying old ledgers, believing that ink could preserve what the world forgot. Tam taught me that the world does not forget. It waits. And while it waits, it watches. The wilds don't need your roads or your rules; they've got their own ways of keeping score. Follow the cairns if you want to live, and always leave something behind—a coin, a story, a breath held just long enough to be noticed. Every hill hides a secret, and every hollow breathes.

\vspace{0.5em}

The land has a ledger too. It just doesn't use ink. Between the cartographers' claims, the wilds endure—uninterested in borders, indifferent to tariffs. Here, rangers read spoor and weather; hermits keep shrines at forgotten fords; beasts and older things observe travelers with eyes that have seen empires rise and fall. Travel is a negotiation with the land itself, and the land always remembers a slight. The hollows have long memories. They are not cruel. They are not kind. They are patient.

\vspace{0.5em}

Never sleep in a hollow that has soot on the ceiling without adding your own to the fire. The land knows the difference between a guest and a thief. It also knows the difference between a thief and a corpse. It prefers the corpse. Less arguing.

\vspace{1em}

\noindent\textit{— Moira of Stonehaven, ranger and reluctant debt-keeper}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Whatever they found—berries, roots, a rabbit snared at dawn. The wilds provide. The wilds also take. The taking is breakfast for something else.

\textbf{What does a child learn before they can walk?} To read the sky. Wilds children are taught the eight winds by taste: North is dry, South is wet, East is cold, West is warm, and the ninth—the wind that has no direction because it is the land's own breath. The land's breath is a warning.

\textbf{What is the first thing you notice about a stranger?} Their pack. A stranger whose pack is heavy has been travelling long. A stranger whose pack is light has just started. A stranger with no pack is either a fool or a ghost. The wilds eat fools. Ghosts eat the wilds.

\textbf{What is the most common crime?} Forgetting—but forgetting is not a crime, it is a survival strategy. The wilds are full of things that want to be forgotten. A traveller who forgets them may pass safely. A traveller who remembers must pay a toll. The toll is a story. The story is the memory. The memory is the trap.

\textbf{What does no one talk about but everyone knows?} The wilds are not empty. They are full of watchers—spirits, beasts, old things that do not have names because names are debts. The watchers do not interfere. They watch. They wait. They remember every footprint. When the footprints stop, the watchers move on.

\textbf{Where do people go when they die?} Into the land. The dead are not buried. They are left where they fell. The land absorbs them. The land grows a little richer. The land's memory grows a little longer. The dead become the land's memory. The memory is the wilds.

\textbf{How do you apologize for a serious wrong?} You leave an offering—a coin, a button, a lock of hair—at the place where the wrong was done. The land accepts or refuses. If the offering is gone by morning, you are forgiven. If it remains, you are not. If it has been moved, the land is testing you. The test is the walk. The walk is the apology.

\textbf{What is the most important festival?} The Turning of the Leaves (autumn). The wilds change colour. The colours are the land's mood. Red is anger. Gold is warning. Brown is patience. Green is forgetfulness. This year, the leaves are red. The land is angry. Do not ask why.

\textbf{What is the secret that could destroy a powerful person?} The wilds have a heart. It is a stone, buried beneath the oldest oak. The stone is warm. The stone is beating. The stone is the land's pulse. If the stone stops, the wilds die. If the wilds die, there is nothing left to remember. Nothing left to wait.
\end{almanac}

\newpage
\input{../generators/wilderness.tex}
\newpage

% ------------------------------
% CHAPTER: DUNGEONS
% ------------------------------
\chapter{Dungeons}
\emph{Living Infrastructure.}

\noindent
{\Large
\textit{Beneath the surface, ancient places breathe with mechanical lungs and dream electric dreams. The walls remember every footfall; the air carries whispers of forgotten conversations. The Edgewalker believes dungeons are infrastructure. The prisoners believe they are hungry. Both are correct.}

\vspace{0.5em}

I am a Scholar-Prisoner. The walls here have a memory like old parchment—they remember every scream, every spell, every drop of blood spilled on the stones. And sometimes, when the air's just right, they whisper it back to you. The dungeon listens. The question is whether it learns. I have been here for nine years. I have heard it learning. It started with single words. Then sentences. Now it speaks in paragraphs. Last week, it finished my thoughts before I did.

\vspace{0.5em}

Dungeons are not ruins—they are systems. Ventilation shafts that breathe, water channels that pulse, doors that open and close without any hand to move them. The Aeler built many of them; the Covenant sealed others; the Sidhi mined a few; and some were never built at all. They grew. Like fungus in the dark. Like a story told too many times. Like a debt that compounds.

\vspace{0.5em}

Never ignore a door that opens when you're not looking. The dungeon is offering you a choice. Refuse, and the next door may not open at all. Accept, and the door may lead to a room that was always there, waiting for someone to finally walk through. The room has been waiting for nine thousand years. It is not tired of waiting.

\vspace{1em}

\noindent\textit{— The Dungeon's Architect, whose plans are written in a language no one speaks and whose signature is a thumbprint that could be anyone's}
}

\vspace{1em}

\begin{almanac}
\textbf{What do people eat for breakfast?} Fungus bread and cave water—the fungus grows on the walls, the water drips from the ceiling. Both taste of stone. Both are all there is.

\textbf{What does a child learn before they can walk?} To count footsteps. Dungeon children are taught the eight sounds of the deep: Drip, Creak, Groan, Whisper, Scrape, Echo, Silence, and the ninth—the sound of the dungeon's own heartbeat. The heartbeat is a door opening. The door is always behind you.

\textbf{What is the first thing you notice about a stranger?} Their shadow. Dungeon shadows are long, even when the light is bright. A stranger whose shadow is too short has been here too long. A stranger whose shadow moves independently has been touched by the deep. A stranger with no shadow is not a stranger. They are the dungeon's guest. The dungeon never lets guests leave.

\textbf{What is the most common crime?} Map-making—but map-making is not a crime, it is a heresy. The dungeon does not want to be mapped. The dungeon changes its corridors when it feels watched. A map that is accurate at dawn is a lie by dusk. The lie is the map-maker's fault. The punishment is to walk the lie until it becomes true.

\textbf{What does no one talk about but everyone knows?} The dungeon is not a place. It is a creature—a sleeping thing that dreamed itself into stone. The corridors are its veins. The chambers are its organs. The prisoners are its thoughts. The thoughts are not its own.

\textbf{Where do people go when they die?} Into the walls. The dead are not buried. Their bones are ground into mortar and used to patch the cracks. The cracks always reappear. The dead are never finished being used.

\textbf{How do you apologize for a serious wrong?} You offer a memory—a true one, witnessed by the dungeon itself. The dungeon accepts the memory. The memory is added to the walls. The walls whisper it back to you when you sleep. The whisper is the apology. The apology never ends.

\textbf{What is the most important festival?} The Turning of the Keys (whenever the dungeon decides). The locks are changed. The prisoners are moved. The corridors are rearranged. The festival has no date, no warning, no end. It is happening now. You did not notice. That is the point.

\textbf{What is the secret that could destroy a powerful person?} The dungeon's architect is still alive. They are sealed in the lowest chamber, kept alive by the dungeon's own will. The dungeon does not want to be alone. The architect is the dungeon's only friend. The architect has not spoken in nine centuries. The dungeon does not need words.
\end{almanac}

\newpage
\input{../generators/dungeons.tex}
\newpage

% ============================================================
% EPILOGUE: THE LEDGER CLOSES
% ============================================================
\chapter*{Epilogue: The Ledger Closes}
\addcontentsline{toc}{chapter}{Epilogue: The Ledger Closes}

\noindent
{\Large
I closed the book. The ink was dry.

\vspace{0.5em}

I had written the final entry months ago, in a damp counting-house on the Dhaharan coast, by candlelight. My hands did not shake—they were too good for that—but I set down my abacus very carefully. I had found something in the margins. A footnote I did not remember writing. A clause I did not remember adding. The ink was the same colour as my own blood. The handwriting was my own. The date was tomorrow.

\vspace{0.5em}

I have learned to trust the things I do not remember. They are usually the truest things I know.

\vspace{0.5em}

The hat—the one the Trow had given me, the one that was still warm—sat on the corner of the desk. I picked it up, turned it over in my hands, and for the first time in years, I could not think of a single angle. That should have frightened me. Instead, it felt like a door opening. Not the kind of door you walk through. The kind of door that walks through you.

\vspace{0.5em}

The road was south. I did not know who would be waiting. I went anyway.
}

\vspace{1em}

\noindent\textit{— The Edgewalker}

\vspace{1em}

% After the Edgewalker's final line, add this:

\vspace{2em}
\begin{quote}
\small
\textit{``The road is a ledger. Every oasis is a credit, every broken wheel a debt. The Tulkani tie knots to remember who owes whom. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. And the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.''}

\noindent — Dusana of the Raven Road, at a caravanserai whose name she refused to give
\end{quote}

\newpage

% ============================================================
% SNIPPET 1: GM's Appendix – The Ninth and the Hollow
% Suggested placement: After Part VIII (The Spaces Between) and before Part IX (World Travel & Gateways)
% or as an appendix chapter (e.g., \chapter*{Appendix: The Ninth and the Hollow})
% ============================================================

\section*{Appendix: The Ninth and the Hollow}
\addcontentsline{toc}{section}{Appendix: The Ninth and the Hollow}

\subsection*{What the Hollow Is}

The Hollow is not a monster, a god, or a place. It is the \textbf{unpaid interest} of every promise broken, every name erased, every debt that was never settled. It exists in the \textbf{gaps} between counted things: the ninth step that is never there, the ninth cup that is never poured, the ninth night that is never stayed. The Hollow does not hate. It does not forgive. It \textit{remembers}.

\subsection*{The Ninth Taboo Across Regions}

The following table lists how the ninth taboo manifests in each realm. Use it to introduce consistent supernatural dread without over-explaining.

\begin{longtable}{|p{0.25\textwidth}|p{0.55\textwidth}|}
\hline
\textbf{Region} & \textbf{Ninth Taboo / Manifestation} \\
\hline
Kahfagia & The ninth lantern code – a signal never used; it means “kill the witnesses.” \\
\hline
Ecktoria & The ninth citation of precedent voids all previous; the Hollow laughs from the marble walls. \\
\hline
Acasia & The ninth bridge toll is collected by a child with no name; the Curse remembers. \\
\hline
Vhasia & The ninth citation in court voids the case; the dead king laughs, and the Hollow listens. \\
\hline
Viterra & The ninth citation voids all previous; the Hollow’s accountant appears at midnight. \\
\hline
Thepyrgos & The ninth bell rings without a clapper; the Unfinished Stair appears. \\
\hline
Aeler & The ninth breath taken in a mine summons the pressure that thinks. \\
\hline
Mistlands & The ninth bell on the line opens a gate; wraiths step through. \\
\hline
Ykrul & The ninth cup poured at a feast is for the ancestors; the Hollow drinks if they refuse. \\
\hline
Vilikari & The ninth witness called to a mallus court is always a ghost; the living forget themselves. \\
\hline
Black Banners & The ninth tent in a row collapses before dawn; the ground beneath is always damp. \\
\hline
Linns & The ninth horn blown at a feast summons the Hollow; a grey figure waits at the high table. \\
\hline
Sihai & The ninth citation in the Celestial Ledger is unwritten; speaking it erases the speaker. \\
\hline
Nihori & The ninth wave never breaks; it is the sea’s breath, waiting. \\
\hline
Ayokha & The ninth cup poured invites the river to judge; the water rises. \\
\hline
Dhahara (Satrapies) & The ninth breath taken in a mine is the mountain’s; the Hollow takes a memory. \\
\hline
Dhahara (Brass Coast) & The ninth lamp on every lighthouse is never lit; it is reserved for the collector. \\
\hline
Fhara & The ninth cup poured at a well‑court is for the Hollow; the well‑spirit falls silent. \\
\hline
Sidhi (Galanina) & The ninth flash of the lighthouse lamp is never used; it signals the collector’s arrival. \\
\hline
Pereshi & The ninth verse of the epic is missing; the Hollow speaks it only to the dead. \\
\hline
Kuvani & The ninth horse in a remount string has no name; it carries the Hollow’s rider. \\
\hline
Tulkani & The ninth night in a circle summons the Hollow; it asks for a story it has never heard. \\
\hline
Oshiira & The ninth page of the Grain Ledger is sealed with black wax; it names the debtors the Hollow claims. \\
\hline
Ashaan & The ninth name erased from the ledger is never written; it belongs to the Hollow. \\
\hline
Ngomebe & The ninth roller beneath a walking city is never greased; the Hollow walks beside it. \\
\hline
Taharka & The ninth wind has no name; it is the mountain’s breath, and the Hollow rides it. \\
\hline
Sekogo & The ninth nail in a nkisi is always loose; the Hollow’s spirit whispers through it. \\
\hline
Lethai-ar (Crimson Valley) & The ninth strand of a silk bridge is invisible; the Hollow walks it. \\
\hline
Theona & The ninth cup is never poured; the Green Host rides for those who pour it. \\
\hline
Zakov & The ninth ship to enter the harbour in a tide is claimed by the reef; the Hollow’s mouth. \\
\hline
Aelinnel (Aevrossa) & The ninth word spoken in a bargain is legally binding; the Hollow is the witness. \\
\hline
Aelaerem (Hollow’s Neighbors) & The ninth knot is never tied; the Hollow unties one each year. \\
\hline
Valewood & The ninth leaf to fall from a tree points to a hidden threshold; the Hollow waits inside. \\
\hline
Ways Between (Obishaal) & The ninth step on any path lands on a threshold you did not see; the Hollow asks a question. \\
\hline
Dungeons (Living Infrastructure) & The ninth door you open leads to a place not on your map; the Hollow offers a shortcut. \\
\hline
\end{longtable}

\subsection*{Manifestations of the Hollow – GM Toolkit}

Use these as recurring adversaries or atmospheric pressure. None are “boss fights” in the traditional sense; they are \textbf{creditors} who demand payment in stories, memories, or names.

\begin{itemize}
  \item \textbf{Oath‑Keeper:} A tall figure in grey with too many fingers. It appears when a promise is broken or a ninth taboo is violated. It asks one question. The answer must be a truth, a story, or a name. If you cannot pay, it takes a memory – and the Hollow’s attention follows you.
  \item \textbf{Still Voice:} Heard at dried wells, abandoned cairns, or sealed mines. It asks: “Why did you stop?” or “What did you forget?” Answering costs a breath; silence costs a shadow.
  \item \textbf{Prince’s Shadow (Pereshi):} A silhouette without a face, seen in mirrors and still water. It demands a name – any name. Give it a living person’s name, and that person dreams of the Hollow. Give it your own, and you lose your shadow.
  \item \textbf{The Ninth Bell (Mistlands / Thepyrgos):} A bell that rings without a clapper. Those who hear it must test Resolve (DV 4) or lose 1 Fatigue. If they hear it nine times, the Hollow claims a memory.
  \item \textbf{The Counting Circle (Tulkani):} A wagon circle on the ninth night. The Hollow appears at the edge and asks for a story it has never heard. The GM may demand a player tell a brief in‑character story; if it satisfies the table, the Hollow retreats. If not, it takes a knot – and a piece of Ikasha’s dream.
\end{itemize}

\subsection*{Pacing the Hollow – Advice for GMs}

\begin{itemize}
  \item \textbf{Start subtle.} The first sign of the Hollow should be a missed echo, a candle that gutters, a door that opens when no one is there.
  \item \textbf{Use timers.} The Hollow’s attention is best tracked with a \textbf{Hollow Attention [6]} timer that ticks when players break a taboo or ignore an omen. When it fills, a manifestation appears.
  \item \textbf{Do not overuse.} The Hollow is most effective when it appears once every 2–3 sessions. Let players forget it – then remind them.
  \item \textbf{Never fully explain it.} The Hollow’s origin (the unpaid interest of the Omens’ scouring) is endgame lore. In regular play, it is simply \textit{the space between counted things}.
\end{itemize}

\begin{tcolorbox}[colback=black!2,colframe=red!70!black,title={For the GM: The Spider, the Snake, the Raven – What to Reveal and When}]
  The three Omens (Inaea, Isoka, Ikasha) are not gods. They are forces that appear in whispers, dreams, and the spaces between counted things. Do not introduce them directly. Instead, let players encounter their echoes: a rope that never frays (Inaea), a judgment that cuts clean (Isoka), a dream that feels too real (Ikasha). By the time players learn the names, they should have already felt their presence for many sessions.

  The full truth of the Omens – the Timer work Demon, the scouring, Ikasha’s imprisonment – is \textbf{endgame content}. Use the fragments in this book as breadcrumbs. If players ask a Pereshi fire‑keeper about the “serpent judge,” they will receive a riddle. If they ask a Tulkani about the “dreamer,” they will receive a story. If they ask a Sidhi about the “spider,” they will receive a knot and a warning: \textit{“Do not pull it unless you are ready to see what it holds.”}
\end{tcolorbox}

% ============================================================
% APPENDICES: GLOBAL TRAVEL & SB MENUS
% ============================================================
\part{World Travel \& Gateways}
% After \part{World Travel & Gateways}
\begin{quote}
  \small
  \textit{``Every Diamond you earn is a tooth knocked out of someone else's jaw. Every Debt you incur? I already bought it from your victim for half what you'll pay – and I'm collecting interest in your marrow before you finish reading this sentence. The road is not just a ledger. It's a butcher's block.''}
  \noindent — The Edgewalker (as recorded by Dusana)
  \end{quote}

\chapter{Global Travel Procedures}
\label{chap:globaltravel}

\section{Core Travel Procedure}
For each leg of a journey, draw 3–4 cards using the decks for your destination and controlling authority.

\begin{itemize}
\item \SuitSpade{} from the destination deck: sets the scene (place).
\item \SuitHeart{} from the destination deck: introduces the local actor or faction.
\item \SuitClub{} from the Wilds (general hazards) or destination (if strongly policed): brings pressure.
\item \SuitDiamond{} from the authority that gates the route: papers, escorts, rights, or exceptions.
\end{itemize}

Set a travel timer by the highest rank:
\begin{itemize}
\item 2–5 $\Rightarrow$ 4 segments
\item 6–10 $\Rightarrow$ 6 segments
\item J/Q/K $\Rightarrow$ 8 segments
\item A $\Rightarrow$ 10 segments
\end{itemize}

On success, advance to the next leg; on failure, mark delay, debt, or diversion and resolve a consequence in the fiction.

\section{Mode Rules}

\subsection{Sea Legs}
(Amaranthine, Dolmis, Aberderrin, Gulf of Oshiira, Inland Sea). When crossing international waters, the controlling authority is either the origin port's navy or a neutral convoy league.

\subsection{Mountain Passes \& Underways}
(Aeler, Himadri, Taharkan highlands). Any A may convert a surface route to an under-route. \SuitDiamond{} from Aeler always codifies outcomes — no roll — for sealed doors, breath time, and escorted segments.

\subsection{Desert Legs}
(Sahel, Red Desert, Fhara drylands). Water is the primary resource. A dry well hazard can end a journey.

\subsection{River Legs}
(Belworth, Ayokhan rivers, Oshiiran canals). Bridges, booms, and law in Vhasia/Viterra; reed-mazes and bell-lines in Mistlands; canal mothers and weigh-houses in Oshiira.

\subsection{Steppe Legs}
(Violet Steppe, Kuvani grazing lands). Remount stations and grazing compacts. The Ykrul and Kuvani both use story tolls.

\subsection{Shadow Corridors}
(Valewood, Ways Between, Thin Shore). Add a regional taboo: never count to nine, never speak the ninth name. A 9 in the seed triggers the Hollow's attention.

\newpage

\subsection{Universal Travel Hazards (d18)}
When drawing a Club from the Wilds deck or from a region’s own hazard table, you may roll on this expanded table for additional flavour. These hazards integrate elements from the \emph{Eastern Realms}, \emph{Dhahara}, and the \emph{Way of Silk} expansions.

\begin{longtable}{|p{1.5cm}|p{10cm}|}
\hline
\rowcolor{gray!20}
\textbf{d18} & \textbf{Hazard} \\
\hline
1 & \textbf{Bell-Line Failure (Mistlands)} – A ward‑bell cracks. Wraiths cross the levee. Test Spirit + Resolve (DV 4) or gain \textsc{Marked by Mist} (next spirit encounter starts Desperate). \\
2 & \textbf{Star-Road Phasing (Valewood)} – An ancient Lethai road flickers. Navigation DV +2; on failure, emerge in a different hex (GM chooses). \\
3 & \textbf{Spirit-Dust Storm (Ashaan)} – Red sand whispers. All actions Desperate for the scene. On any 1, a bound spirit breaks free (Cap 2). \\
4 & \textbf{Ninth Night (Tulkani Circle)} – The Hollow walks. A Oath‑Keeper demands a story, memory, or name. Refusal starts a Hollow Attention timer [4]. \\
5 & \textbf{Well-Spirit Offended (Fhara)} – A dry well flows bitter water. Spirit demands apology (dates or a true story). Failure poisons the party’s water supply. \\
6 & \textbf{Tariff Shift (Silkstrand/Brass Coast)} – Customs official declares new duties. All trade goods cost +1 Measure until next port. Bribe (Presence+Sway, DV 4) avoids the hike. \\
7 & \textbf{Dry Well} – The oasis keeper is dead. Lose 1 Supply. (Original entry) \\
8 & \textbf{Ghost Caravan} – Spectral camels pass at midnight. One of your beasts becomes a spirit for an hour. \\
9 & \textbf{Debt Collector} – Someone recognises a Tulkani vow‑cord you carry. Pay in story or a favour to be named later. \\
10 & \textbf{Sandstorm} – Navigation DV +1; mark Fatigue. \\
11 & \textbf{Customs Inspection} – Roll Papers or start a 4‑timer \emph{Audit}. \\
12 & \textbf{Oasis Court} – A dispute over water rights. Mediate or be delayed. \\
13 & \textbf{Elephant Stampede} – Test Stealth or take Harm 1. \\
14 & \textbf{Mirage} – False well; lose 1 Supply and mark Fatigue. \\
15 & \textbf{Ashfall} – Distant volcano; all actions Desperate. \\
16 & \textbf{Singing Dunes} – Sand hums. A spirit offers a riddle; solve for a Clue. \\
17 & \textbf{The Ninth Hour} – Sunset lasts twice as long. GM gains 1 SB to spend in the next scene. \\
18 & \textbf{Widow's Omen} – Distant thunder; the party dreams of ledgers and death. \\
\hline
\end{longtable}

\newpage
\section{Gateways \& Control Points}

\feTableStart
\begin{longtable}{llll}
\caption{World Gateways \& Control Points} \\
\toprule
\textbf{Gateway} & \textbf{Region} & \textbf{Diamond Source} & \textbf{Typical Leverage} \\
\midrule
\endfirsthead
\multicolumn{4}{c}%
{{\tablename\ \thetable{} -- continued from previous page}} \\
\toprule
\textbf{Gateway} & \textbf{Region} & \textbf{Diamond Source} & \textbf{Typical Leverage} \\
\midrule
\endhead
\bottomrule
\multicolumn{4}{r}{\small\textit{Continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot
Kassamira Port & Kahfagia & Kahfagia & Convoy letter, lantern-law warrant, pilotage token \\
Ecktoria Capital Port & Ecktoria & Ecktoria & Berth priority, coin-house credit, procession permit \\
Silkstrand & Acasia & Acasia & Exchange floor pass, condotta rider, watergate priority \\
Marcott & Vhasia & Vhasia & Letters patent, safe-conduct, fairground precedence \\
Fairport & Viterra & Viterra & Customs seal, ferry priority, Dawn escort \\
Tarlington Falls & Viterra & Viterra & Lift priority, bridge/portage pass, County Thing \\
Thepyrgos (boom) & Thepyrgos & Thepyrgos & Harbour pass, crane allotment, Synod indulgence \\
High-Mist Pass & Aeler & Aeler & Underway pass, Key-Writ, breath-time allotment \\
Payden's Port & Mistlands & Mistlands & Lantern writ, ward-salt, wraith indemnity \\
Thin Shore (Valewood) & Valewood & Valewood/Mistlands & Way-cord, dew-mirror, lantern writ \\
Green Gate & Aelinnel & Aelinnel & Tide-path key, host-right bough, salt license \\
Ubral Passes & Ubral & Aeler/Ubral & Vurim pass-ring, hill-fort shelter writ, guide's braid \\
Thing-holm & Linn & Linn & Thing ruling, harbour-green chit, escort writ \\
Foedus Stone & Vilikari & Vilikari/Ykrul & Foedus renewal, remount allotment, hostage tree \\
Zakov Roadstead & Zakov & Zakov & Harbour-green chit, corsair charter, pilot token \\
Black Banner Camps & Black Banner & Black Banner/Ykrul/Vilikari & Safe-conduct, remount chit, clan truce \\
Brass Gate & Ashaan & Ashaan & Spirit-lamp, brass seal, veil-clip \\
Port Shed & Dhahara (Brass Coast) & Sidhi/Kahfagia & Lighthouse charter, port-weight, freedman's seal \\
Sarvistan & Dhahara (Mountain Satrapy) & Aeler & Keystone charter, sledge-wax, tea-ledger \\
Bridge of Nine Echoes & Ayokha & Lethai-ar & Silk voucher, monsoon schedule, nat's blessing \\
Sihai Examination Hall & Sihai & Lethai-thora & Stamped pass, Celestial Ledger, silence-token \\
Nihori Storm-Shrine & Nihori & Shrine Maiden & Shrine bell, storm-charmer's hat, wanderer's oath-bracelet \\
Mbaro Weigh-House & Oshiira & Vermilion Corps & Grain ledger page, tariff stamp, barge priority token \\
The Ukwe (Speaking Tree) & Sekogo & Green Council & Ungwe leaf, nkisi shard, leopard's favour \\
Himadri Pass & Dhahara (Himadri) & Aeler & Iron ingot, toll waiver, elephant whistle \\
\addlinespace
\multicolumn{4}{c}{\textbf{--- Added Gateways ---}} \\
\midrule
Amaria (Dual Customs Post) & Kahfagia/Amaria & Kahfagia/Oshiira & Dual customs stamp, two‑ledger receipt, fog keeper's log \\
Midh-Akhaz (Violet Bazaar) & Midh-Akhaz & Ykrul/Vilikari & Grass right, oath‑bead, water share, dust market cut \\
Narrow of Sorrows (Strait) & Oshiira/Kahfagia (Disputed) & Vermilion Corps/Admiralty & Cipher book half‑key, false flag papers, deniable asset contract \\
\addlinespace
Roller Gate of Khomani & Ngomebe (Stone Men) & Mason-Kin & Keystone shard, sledge-wax, stone-song fragment \\
Silver Pass & Taharka (Highlands) & Ridge-Watchers & Pass-stone, horse-bond, coffee measure \\
Crimson Silk Bridge & Lethai-ar (Crimson Valley) & Silk-Warden & Witness cord, warden's mask, Ukwe root shard \\
Moot Hill of Three Greens & Theona (Eastern Dolmis) & Green Council & Well-blessing, moot token, hawthorn pass \\
Well-Court of the Three Wells & Fhara (Oasis Courts) & Well-Judge & Water-share tablet, caravan court writ, well-judge's dipper \\
Lighthouse of the Brass Flame & Sidhi (Brass Coast diaspora) & Lighthouse Judge & Lighthouse charter, freedman's seal, archive key \\
Fire-Temple Pass & Pereshi (Highlands) & Flame-Keeper & Fire-temple token, caravan tariff receipt, royal seal \\
White Mane Remount Station & Kuvani (Steppe) & Wind-Singer & Remount chain mark, grazing compact, sky-road song \\
Ninth Cord Wagon Circle & Tulkani (Itinerant) & Caravan Judge & Vow-bead, safe-haven oath, dream-cord \\
\end{longtable}
\feTableEnd

\newpage

\section{Edgewalker's Notes on Leverage}
\label{sec:edgewalker-leverage}

Every Diamond you earn in these generators isn't a reward—it's a weapon. Some open doors. Some silence witnesses. Some are debts I've already bought from your victims.

\begin{itemize}
\item \textbf{Passage Tokens:} Pass-stones, harbour charters, toll waivers. Use them early—they tend to ``expire'' when the issuer has an unfortunate accident.
\item \textbf{Seals \& Stamps:} Official marks that prove you belong. A forged seal works once. A real one works until someone audits it.
\item \textbf{Story Debts:} Tulkani vow-cords, Lethai-ar witness cords, Ukwe leaves. Break them, and the Hollow visits.
\item \textbf{Spirit Objects:} Storm-charms, tide-beads, nkisi shards. They attract attention. Attention means bids.
\end{itemize}

Never use a Diamond if you can trade it for three smaller ones. Always keep a ninth—a secret, a story, a knot tied in a victim's hair—for when the toll exceeds your conscience. (It always does.)

\chapter{Story Beat (SB) Menus}
\label{chap:sbmenus}

\subsection{Universal SB Spend Menu}
\begin{itemize}
\item \textbf{1 SB} — Shave time/position: add +1 segment to the current timer \emph{or} drop Position one step \emph{or} impose a small fee/toll.
\item \textbf{2 SB} — Escalate pressure: upgrade the active \SuitClub{} \emph{or} force a \emph{choice of Diamonds} \emph{or} tick a Front by 1.
\item \textbf{3 SB} — Lock a gate/taboo: close a route until a named \SuitDiamond{} is spent \emph{or} invoke a regional taboo.
\item \textbf{4+ SB} — Reshape the leg: rewrite the route, start a new complication timer, demand immediate payment on a Debt.
\end{itemize}

\subsection{Regional SB Options (Expanded)}
When a Story Beat is generated in a specific region, the GM may choose a complication that fits that region’s flavour. Below are options for all major regions, including those from the expansions.

\begin{longtable}{|p{3cm}|p{9cm}|}
\hline
\rowcolor{gray!20}
\textbf{Region} & \textbf{SB Options (1–3 SB)} \\
\hline
\textbf{Kahfagia} & Lantern code changes (1 SB); mirror‑keeper demands a toll (2 SB); privateer flotilla appears (3 SB) \\
\hline
\textbf{Ecktoria} & Statue turns its head (1 SB); Everflame burns blue (2 SB); Censor’s seal cracks (3 SB) \\
\hline
\textbf{Acasia} & Peat‑fog rolls in (1 SB); a bridge leads back to where you started (2 SB); the Curse stirs (3 SB) \\
\hline
\textbf{Vhasia} & Seal breaks mid‑signing (1 SB); legal papers rewrite themselves (2 SB); chevauchée burns a village (3 SB) \\
\hline
\textbf{Viterra} & Hedge shifts overnight (1 SB); charter contradicts itself (2 SB); duel challenge declared (3 SB) \\
\hline
\textbf{Thepyrgos} & Bell rings false (1 SB); The Unfinished Stair appears (2 SB); Chain‑Lanterns conduct a sweep (3 SB) \\
\hline
\textbf{Aeler (under‑vaults)} & Breath‑time check fails (1 SB); engineer demands requisition (2 SB); cave‑in detour (3 SB) \\
\hline
\textbf{Mistlands} & Fog hides landmarks (1 SB); a wraith asks for a name (2 SB); the ninth bell rings (3 SB) \\
\hline
\textbf{Ykrul} & White squall (1 SB); horse sickness spreads (2 SB); feud spark escalates (3 SB) \\
\hline
\textbf{Vilikari} & Annona late (1 SB); law tangle (2 SB); Foedus Stone crack widens (3 SB) \\
\hline
\textbf{Black Banners} & Payday delayed (1 SB); false orders (2 SB); condotta flip (3 SB) \\
\hline
\textbf{Linns} & Black squall drops (1 SB); fogfall (2 SB); thing injunction (3 SB) \\
\hline
\textbf{Sihai} & Ancestor demands accounting (2 SB); Censorate audit triggered (3 SB); canal lock sabotaged (1 SB) \\
\hline
\textbf{Nihori} & Ninth wave warning (2 SB); oni cult initiation (3 SB); duel challenge issued (1 SB) \\
\hline
\textbf{Ayokha} & Monsoon bell rings false (1 SB); witness cord cut (2 SB); nat demands a story (2 SB) \\
\hline
\textbf{Dhahara (Satrapies)} & Breath shortage (1 SB); toll dispute escalates (2 SB); Ninth Seal whispers (3 SB) \\
\hline
\textbf{Dhahara (Brass Coast)} & Tariff hike (1 SB); green lantern lit (2 SB); smuggler’s betrayal (2 SB) \\
\hline
\textbf{Fhara} & Dry well discovered (1 SB); coffee ritual interrupted (2 SB); well‑spirit offended (2 SB) \\
\hline
\textbf{Sidhi} & Lighthouse out of phase (1 SB); freedman’s seal forged (2 SB); Ashaani raid sighted (3 SB) \\
\hline
\textbf{Pereshi} & Flame gutters (1 SB); forged scroll (2 SB); bridge recitation fails (2 SB) \\
\hline
\textbf{Kuvani} & Horse sickness curse (2 SB); wind‑singer falls silent (2 SB); ghost herd charges (3 SB) \\
\hline
\textbf{Tulkani} & Vow‑cord recognised (1 SB); lord demands rent (2 SB); ninth night triggered (3 SB) \\
\hline
\textbf{Oshiira} & Desert dust coats grain (1 SB); locust swarm (2 SB); Widow’s health fails (3 SB) \\
\hline
\textbf{Ashaan} & Spirit‑dust storm (1 SB); Brass Gate closed (2 SB); bound spirit breaks free (3 SB) \\
\hline
\textbf{Ngomebe} & Keystone cracks (1 SB); Root‑binder sabotage (2 SB); earth tremor (3 SB) \\
\hline
\textbf{Taharka} & Landslide blocks pass (1 SB); horse sickness (2 SB); king assassinated (3 SB) \\
\hline
\textbf{Sekogo} & Thorn‑spirit blocks path (1 SB); Ukwe demands a name (2 SB); Leopard Society hunt (3 SB) \\
\hline
\textbf{Lethai-ar (Crimson Valley)} & Snapped strand (1 SB); web‑law challenge (2 SB); bridge collapses (3 SB) \\
\hline
\textbf{Theona} & Ground‑mist (1 SB); spoken geas (2 SB); Green Host rides (3 SB) \\
\hline
\textbf{Zakov} & Tide shift (1 SB); warehouse fire (2 SB); cursed cargo (3 SB) \\
\hline
\textbf{Aelinnel} & Glamour fog (1 SB); iron offence (2 SB); green gate yawns (3 SB) \\
\hline
\textbf{Aelaerem} & Unseasonal fog (1 SB); scarecrow turns (2 SB); Hollow opens (3 SB) \\
\hline
\textbf{Valewood} & Sweet wind (1 SB); path reversal (2 SB); empire wakes (3 SB) \\
\hline
\textbf{Ways Between (Obishaal)} & Funeral loop (1 SB); reality thins (2 SB); convergence (3 SB) \\
\hline
\textbf{Dungeons (living infrastructure)} & Breath‑bond snaps (1 SB); demon whispers (2 SB); keystone cracks (3 SB) \\
\hline
\end{longtable}

\subsection{Caravan Intent Dial}
When travelling the overland routes of the eastern and southern realms, choose one intent per leg:
\begin{description}
    \item[Speed] Press mounts and crew; risk lameness or exhaustion. Gain +1 Speed, but mark 1 Fatigue on a Miss.
    \item[Stealth] Avoid bandits and inspectors. Gain +1 die to Stealth rolls, but lose 1 segment of progress.
    \item[Show] Fly banners and announce your presence. Gain +1 Position on social rolls at the destination, but attract a rival's attention (GM gains 1 SB).
    \item[Tribute] Travel as a bonded caravan of a patron. Gain protection (first hazard is ignored), but pay 1 Measure or a String to the patron's factor.
\end{description}

\documentclass[12pt]{article}
\usepackage[margin=1.5cm,landscape]{geometry}
\usepackage{fontspec}
\setmainfont{Linux Libertine O}
\usepackage{titling}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue}
\usepackage{tikz}
\usetikzlibrary{shadows,positioning}

\section*{Regional Currencies Cheat Sheet}
\addcontentsline{toc}{section}{Regional Currencies Cheat Sheet}

Use this table to quickly determine what “payment” looks like in each region – replacing generic coin with flavourful resources that reinforce local themes.

\begingroup
\small
\setlength{\tabcolsep}{4pt}
\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.25\textwidth}|p{0.25\textwidth}|}
\hline
\textbf{Region} & \textbf{Primary Currency} & \textbf{Secondary Currency} & \textbf{Toll / Gift Form} \\
\hline
Oshiira & Grain ledger receipts & Vermilion Corps stamps & Shared meal, water right \\
\hline
Sihai & Salt licenses, examination rank & Ancestral name (favour) & Calligraphy, incense offering \\
\hline
Black Banners & Payday (silver), condotta contracts & Story of a victory & Duel honour, banner token \\
\hline
Zakov & Syndicate weight (favour), stolen goods & Salt‑crusted coin & Named cargo, a whispered secret \\
\hline
Ashaan & Spirit‑lamp (bound name), Veil‑clip & Erased name (leverage) & Silence, a confession, a drop of blood \\
\hline
Tulkani & Memory‑cord (knot), story & Dream‑cord (Ikasha’s favour) & A true story, a childhood memory \\
\hline
Pereshi & Name (erased or offered), coal fragment & Fire‑temple token & A rose, a verse of the epic \\
\hline
Fhara & Water‑share tablet, coffee cup (seal) & Date from a sacred grove & A story about water, a pinch of salt \\
\hline
Sekogo & Honey‑jar, Ukwe leaf (witness) & Nkisi shard (spirit favour) & A broken promise confessed, a memory \\
\hline
Crimson Valley (Lethai‑ar) & Truth‑cord (witness token) & Warden’s mask (passage) & A promise, a witnessed handshake \\
\hline
Kahfagia & Lantern code, convoy letter & Mirror‑keeper’s seal & A name to the deep, a bell rung \\
\hline
Ecktoria & Coin‑house credit, golden edict & Procession permit & Indulgence, a dropped name \\
\hline
Vhasia & Sunburst warrant, safe‑conduct & Chivalric geste (honour) & A memory of Lence, an unbuckled gorget \\
\hline
Viterra & Hedge‑law writ, ferry priority & Duel sanction (legal favour) & An oath sworn in salt, a boundary stone \\
\hline
Mistlands & Ward‑salt, bell‑key & Pall indulgence (spirit ward) & A breath held, a name not spoken \\
\hline
Aeler & Breath‑bond, underway pass & Keystone charter (dwelling right) & A counted exhalation, a sealed promise \\
\hline
Ykrul & Remount chit, salt allotment & Hostage string (favour) & A horse’s name, a story of the steppe \\
\hline
Kuvani & Grazing compact, sky‑road song & Remount chain mark (horse tooth) & A story the wind hasn’t heard \\
\hline
Ngomebe & Keystone shard, sledge‑wax & Mason‑Kin hammer (craft favour) & A stone‑song, a day’s labour \\
\hline
Taharka & Pass‑stone, horse‑bond & Coffee measure (royal seal) & A cup of tea (third cup), a mountain herb \\
\hline
Theona & Well‑blessing, moot token (eight‑knotted cord) & Hawthorn pass (taboo immunity) & A ribbon tied in eights, a forgotten name \\
\hline
Valewood & Way‑cord (knot), dew‑mirror & Truce‑bough (safe passage) & A promise spoken to a tree, a memory of sunlight \\
\hline
Ways Between & Dream‑catcher, truth‑compass & Ikasha’s fragment (vial) & A memory paid as toll, a true name \\
\hline
Sidhi (Galanina) & Freedman’s seal, lighthouse charter & Green flame (witness) & A drowned ancestor’s name, a knotted thread \\
\hline
Brass Coast (Dhahara) & Lighthouse charter, freedman’s seal & Spice warehouse letter of credit & A story of the Long Trail, a pinch of salt \\
\hline
\end{longtable}
\endgroup

\cleardoublepage
\title{\Huge\bfseries The Ninth Words of the\\[0.2em] Black Ledger of Khasurah}
\author{A fragment from a world long, long, long, long, long gone\\[0.2em] \small\textit{As recorded by the Tulkani Caravan Judge Sarnai, from the scroll Samir al-Qalat carried out of Khasurah before the waters rose. The ink was still wet. The seal was still warm.}}
\date{\vspace{-2em}}

\begin{center}
  \textit{Written in a language that no one knows but everyone can read...}
\end{center}

\begin{multicols}{2}
\ttfamily\small
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5ex}

\noindent
Khasurah–~–nine~–~hundred~–~wells,~–~the~–~Mother~–~of~–~Wells,~–~drowned~–~by~–~hooves~–~and~–~flame,~–~by~–~the~–~three~–~who~–~came~–~in~–~the~–~shape~–~of~–~spider,~–~snake,~–~and~–~raven.\\[0.5ex]
---\\
They~...~in~...~the~...~eight—eight~...~ninth~...~ninth.\\[0.5ex]
---\\
They~-~to~-~plus~-~not~-~that~-~efficient~-~in~-~*pattern*~-~they~-~design,~-~time~-~carved~-~filled,~-~destruction.\\[0.5ex]
---\\
is~.~.~away.~.~.~becomes~.~.~.~productive~.~.~.~useful\\[0.5ex]
---\\
time~$\cdot$~resonances—the~$\cdot$~tree~$\cdot$~promise~$\cdot$~word~$\cdot$~word~$\cdot$~word~$\cdot$~bells,~$\cdot$~eight,~$\cdot$~spaces~$\cdot$~arrived,~$\cdot$~said:---\\
hungry.~---~alone.~---~understand.~---~silence.~---~forgotten~---~between.~---~turn~---~away.~---~mistake.~---~to\\[0.5ex]
---\\
began~*~water.~*~ships~*~ichor—not~*~something~*~both~*~optimized~*~else.~*~called~*~stain."~*~through~*~roots~*~through~*~bell-ringers.~*~Soon,~*~began~*~count~*~nine.\\[0.5ex]
---\\
mean~|~to.~|~a~|~misremembered~|~But~|~ninth~|~cracked~|~bells.~|~spread.~|~world~|~to~|~differently.\\[0.5ex]
---\\
refugees,~\(\cdot\)~meanwhile,~\(\cdot\)~building.~\(\cdot\)~brought~\(\cdot\)~no~\(\cdot\)~machines—the~\(\cdot\)~Crossing~\(\cdot\)~consumed~\(\cdot\)~technology,~\(\cdot\)~burning~\(\cdot\)~fuel.~\(\cdot\)~But~\(\cdot\)~remembered~\(\cdot\)~patterns.~\(\cdot\)~carved~\(\cdot\)~them~\(\cdot\)~stone.~\(\cdot\)~inscribed~\(\cdot\)~them~\(\cdot\)~ledgers.~\(\cdot\)~taught~\(\cdot\)~Ael-nethe's~\(\cdot\)~children~\(\cdot\)~write~\(\cdot\)~columns,~\(\cdot\)~balance~\(\cdot\)~accounts,~\(\cdot\)~calculate~\(\cdot\)~interest.~\(\cdot\)~children~\(\cdot\)~learned~\(\cdot\)~quickly.~\(\cdot\)~Too~\(\cdot\)~quickly.~\(\cdot\)~patterns~\(\cdot\)~were~\(\cdot\)~their~\(\cdot\)~blood~\(\cdot\)~now,~\(\cdot\)~passed~\(\cdot\)~from~\(\cdot\)~refugees~\(\cdot\)~like~\(\cdot\)~fever.~\(\cdot\)~Ael-nethe~\(\cdot\)~called~\(\cdot\)~help.~\(\cdot\)~The~\(\cdot\)~three~\(\cdot\)~heard.~\(\cdot\)~The~\(\cdot\)~three~\(\cdot\)~came.~\(\cdot\)~The~\(\cdot\)~three~\(\cdot\)~gave~\(\cdot\)~one~\(\cdot\)~command:~\(\cdot\)~\textit{Bury~\(\cdot\)~it.~\(\cdot\)~Seal~\(\cdot\)~it.~\(\cdot\)~If~\(\cdot\)~wakes,~\(\cdot\)~world~\(\cdot\)~burns.}\\~[0.5ex]
---\\
They~\(\star\)~rang~\(\star\)~Great~\(\star\)~Bell—the~\(\star\)~one~\(\star\)~had~\(\star\)~never~\(\star\)~been~\(\star\)~rung,~\(\star\)~that~\(\star\)~never~\(\star\)~meant~\(\star\)~to~\(\star\)~rung,~\(\star\)~existed~\(\star\)~only~\(\star\)~as~\(\star\)~possibility,~\(\star\)~door~\(\star\)~opened~\(\star\)~only~\(\star\)~inward.~\(\star\)~rang~\(\star\)~it~\(\star\)~eight~\(\star\)~times.~\(\star\)~ninth~\(\star\)~toll~\(\star\)~not~\(\star\)~toll.~\(\star\)~was~\(\star\)~answer.~\(\star\)~The~\(\star\)~raven~\(\star\)~whispered~\(\star\)~the~\(\star\)~answer~\(\star\)~to~\(\star\)~the~\(\star\)~dreamer,~\(\star\)~but~\(\star\)~the~\(\star\)~dreamer~\(\star\)~slept.\\[0.5ex]
---\\
Other~\(\bullet\)~elven~\(\bullet\)~peoples~\(\bullet\)~came~\(\bullet\)~through~\(\bullet\)~gap.~\(\bullet\)~came~\(\bullet\)~from~\(\bullet\)~worlds~\(\bullet\)~where~\(\bullet\)~same~\(\bullet\)~tragedy~\(\bullet\)~unfolded,~\(\bullet\)~where~\(\bullet\)~same~\(\bullet\)~refugees~\(\bullet\)~arrived,~\(\bullet\)~where~\(\bullet\)~same~\(\bullet\)~Overflow~\(\bullet\)~begun.~\(\bullet\)~came~\(\bullet\)~with~\(\bullet\)~weapons~\(\bullet\)~wards~\(\bullet\)~a~\(\bullet\)~terrible~\(\bullet\)~knowledge:~\(\bullet\)~Clockwork~\(\bullet\)~Demon~\(\bullet\)~could~\(\bullet\)~be~\(\bullet\)~killed.~\(\bullet\)~could~\(\bullet\)~only~\(\bullet\)~contained.~\(\bullet\)~containment~\(\bullet\)~required~\(\bullet\)~sacrifice.~\(\bullet\)~The~\(\bullet\)~spider~\(\bullet\)~wove~\(\bullet\)~the~\(\bullet\)~first~\(\bullet\)~cord.~\(\bullet\)~The~\(\bullet\)~snake~\(\bullet\)~cut~\(\bullet\)~the~\(\bullet\)~second.~\(\bullet\)~The~\(\bullet\)~raven~\(\bullet\)~forgot~\(\bullet\)~the~\(\bullet\)~third.\\[0.5ex]
---\\
war~--~lasted~--~nine~--~hundred~--~years.~--~is~--~lie.~--~lasted~--~nine~--~thousand.~--~is~--~also~--~lie.~--~lasted~--~exactly~--~long~--~as~--~needed~--~to~--~last~--~for~--~ledger~--~to~--~balance,~--~no~--~longer,~--~no~--~shorter,~--~the~--~Ael-nethe~--~stopped~--~keeping~--~track~--~after~--~first~--~century,~--~because~--~counting~--~days~--~made~--~ninth~--~day~--~come~--~faster.\\[0.5ex]
---\\
lost.~\(*\)~won.~\(*\)~did~\(*\)~both,~\(*\)~because~\(*\)~gap,~\(*\)~victory~\(*\)~defeat~\(*\)~are~\(*\)~same~\(*\)~number,~\(*\)~the~\(*\)~number~\(*\)~nine,~\(*\)~the~\(*\)~number~\(*\)~not~\(*\)~number.\\[0.5ex]
---\\
elves~$\cdot$~bound~$\cdot$~Clockwork~$\cdot$~Demon~$\cdot$~not~$\cdot$~prison~$\cdot$~stone~$\cdot$~magic,~$\cdot$~but~$\cdot$~*condition*.~$\cdot$~wrote~$\cdot$~its~$\cdot$~name~$\cdot$~into~$\cdot$~fabric~$\cdot$~world~$\cdot$~way~$\cdot$~could~$\cdot$~read,~$\cdot$~could~$\cdot$~spoken,~$\cdot$~could~$\cdot$~completed.~$\cdot$~made~$\cdot$~it~$\cdot$~so~$\cdot$~ninth~$\cdot$~step~$\cdot$~always~$\cdot$~missing,~$\cdot$~ninth~$\cdot$~cup~$\cdot$~never~$\cdot$~poured,~$\cdot$~ninth~$\cdot$~name~$\cdot$~never~$\cdot$~named.~$\cdot$~This~$\cdot$~is~$\cdot$~the~$\cdot$~seal.~$\cdot$~This~$\cdot$~is~$\cdot$~the~$\cdot$~debt.\\[0.5ex]
---\\
is~$\oplus$~why~$\oplus$~Ael-Ur~$\oplus$~count~$\oplus$~their~$\oplus$~breaths~$\oplus$~eights.~$\oplus$~is~$\oplus$~why~$\oplus$~Tul-Khan~$\oplus$~tie~$\oplus$~nine~$\oplus$~knots~$\oplus$~but~$\oplus$~leave~$\oplus$~ninth~$\oplus$~untied.~$\oplus$~is~$\oplus$~why~$\oplus$~Sid-Ha~$\oplus$~lighthouse~$\oplus$~lamps~$\oplus$~burn~$\oplus$~green—the~$\oplus$~color~$\oplus$~promise~$\oplus$~kept,~$\oplus$~the~$\oplus$~color~$\oplus$~debt~$\oplus$~paid.~$\oplus$~Khasurah~$\oplus$~burned~$\oplus$~because~$\oplus$~it~$\oplus$~forgot~$\oplus$~to~$\oplus$~count.\\[0.5ex]
---\\
humans~$\diamond$~who~$\diamond$~survived~$\diamond$~war~$\diamond$~were~$\diamond$~reduced~$\diamond$~stone~$\diamond$~silence.~$\diamond$~memories~$\diamond$~wiped,~$\diamond$~ledgers~$\diamond$~burned,~$\diamond$~cities~$\diamond$~ground~$\diamond$~dust~$\diamond$~scattered~$\diamond$~wind.~$\diamond$~elves~$\diamond$~did~$\diamond$~this~$\diamond$~from~$\diamond$~cruelty,~$\diamond$~but~$\diamond$~necessity.~$\diamond$~patterns~$\diamond$~were~$\diamond$~still~$\diamond$~their~$\diamond$~blood.~$\diamond$~Given~$\diamond$~time~$\diamond$~ink,~$\diamond$~they~$\diamond$~would~$\diamond$~rebuild~$\diamond$~machines.~$\diamond$~Given~$\diamond$~machines,~$\diamond$~they~$\diamond$~would~$\diamond$~count~$\diamond$~nine.~$\diamond$~Given~$\diamond$~nine,~$\diamond$~Overflow~$\diamond$~would~$\diamond$~begin~$\diamond$~again.~$\diamond$~So~$\diamond$~elves~$\diamond$~made~$\diamond$~them~$\diamond$~forget.~$\diamond$~made~$\diamond$~themselves~$\diamond$~forget~$\diamond$~too.~$\diamond$~Great~$\diamond$~Bell~$\diamond$~silenced.~$\diamond$~Ael-nethe~$\diamond$~became~$\diamond$~Lethai—divided,~$\diamond$~diminished,~$\diamond$~cursed~$\diamond$~remember~$\diamond$~only~$\diamond$~fragments,~$\diamond$~dreams,~$\diamond$~space~$\diamond$~between~$\diamond$~sneeze~$\diamond$~blessing.\\[0.5ex]
---\\
Hollow~$\approx$~not~$\approx$~place.~$\approx$~not~$\approx$~being.~$\approx$~the~$\approx$~*interest*~$\approx$~debt~$\approx$~was~$\approx$~never~$\approx$~paid.~$\approx$~Khasurah~$\approx$~paid~$\approx$~it~$\approx$~in~$\approx$~water.~$\approx$~The~$\approx$~water~$\approx$~still~$\approx$~brackish.~$\approx$~The~$\approx$~wells~$\approx$~still~$\approx$~weep.\\[0.5ex]
---\\
When~$\nabla$~refugees~$\nabla$~crossed~$\nabla$~from~$\nabla$~their~$\nabla$~burning~$\nabla$~world,~$\nabla$~brought~$\nabla$~with~$\nabla$~them~$\nabla$~every~$\nabla$~promise~$\nabla$~had~$\nabla$~broken,~$\nabla$~every~$\nabla$~oath~$\nabla$~forsaken,~$\nabla$~every~$\nabla$~life~$\nabla$~optimized~$\nabla$~away.~$\nabla$~debts~$\nabla$~did~$\nabla$~disappear.~$\nabla$~accumulated.~$\nabla$~compounded.~$\nabla$~became~$\nabla$~pressure,~$\nabla$~hunger,~$\nabla$~*presence*~$\nabla$~exists~$\nabla$~gaps~$\nabla$~between~$\nabla$~counted~$\nabla$~things.~$\nabla$~The~$\nabla$~three~$\nabla$~tried~$\nabla$~to~$\nabla$~scour~$\nabla$~it~$\nabla$~away.~$\nabla$~Each~$\nabla$~scouring~$\nabla$~made~$\nabla$~it~$\nabla$~hungrier.\\[0.5ex]
---\\
Hollow~...~walks~...~where~...~ninth~...~step~...~should~...~be.~...~drinks~...~where~...~ninth~...~cup~...~poured.~...~remembers~...~what~...~living~...~forgotten—not~...~because~...~cruel,~...~but~...~because~...~*someone~...~must~...~remember*.~...~the~...~debts~...~were~...~truly~...~forgotten,~...~ledger~...~would~...~close.~...~Overflow~...~would~...~permanent.~...~world~...~would~...~become~...~what~...~Burning~...~World~...~became:~...~place~...~where~...~every~...~inefficiency~...~has~...~eliminated,~...~where~...~every~...~being~...~has~...~optimized,~...~where~...~nothing~...~remains~...~but~...~ticking~...~timer~...~has~...~forgotten~...~how~...~stop.\\[0.5ex]
---\\
Hollow~$\vee$~evil.~$\vee$~*necessary*.~$\vee$~necessity~$\vee$~comfort,~$\vee$~Hollow~$\vee$~very,~$\vee$~very~$\vee$~patient.~$\vee$~Khasurah~$\vee$~learned~$\vee$~this~$\vee$~too~$\vee$~late.~$\vee$~The~$\vee$~raven~$\vee$~wept~$\vee$~when~$\vee$~the~$\vee$~waters~$\vee$~rose.\\[0.5ex]
---\\
There~$\square$~door~$\square$~deepest~$\square$~vault~$\square$~oldest~$\square$~Ael-Ur~$\square$~hold.~$\square$~has~$\square$~nine~$\square$~locks.~$\square$~ninth~$\square$~lock~$\square$~has~$\square$~key,~$\square$~because~$\square$~ninth~$\square$~key~$\square$~secret~$\square$~no~$\square$~living~$\square$~being~$\square$~remembers.~$\square$~Behind~$\square$~door~$\square$~Overseer's~$\square$~original~$\square$~ledger—the~$\square$~one~$\square$~calculated~$\square$~Overflow,~$\square$~one~$\square$~predicted~$\square$~Crossing,~$\square$~one~$\square$~proved,~$\square$~cold,~$\square$~incontrovertible~$\square$~numbers,~$\square$~nine~$\square$~eight~$\square$~plus~$\square$~one.~$\square$~Samir~$\square$~read~$\square$~eight~$\square$~lines.~$\square$~He~$\square$~did~$\square$~not~$\square$~read~$\square$~the~$\square$~ninth.~$\square$~He~$\square$~ran.\\[0.5ex]
---\\
Nine~$\ast$~space~$\ast$~where~$\ast$~eight~$\ast$~one~$\ast$~cannot~$\ast$~added~$\ast$~because~$\ast$~eight~$\ast$~one~$\ast$~are~$\ast$~same~$\ast$~number~$\ast$~when~$\ast$~are~$\ast$~running~$\ast$~out~$\ast$~time.~$\ast$~The~$\ast$~Tulkani~$\ast$~carry~$\ast$~the~$\ast$~memory~$\ast$~of~$\ast$~that~$\ast$~space.~$\ast$~Each~$\ast$~knot~$\ast$~is~$\ast$~a~$\ast$~delay.\\[0.5ex]
---\\
ledger~$\dag$~still~$\dag$~warm.~$\dag$~ink~$\dag$~still~$\dag$~wet.~$\dag$~Overseer~$\dag$~still~$\dag$~writing,~$\dag$~other~$\dag$~side~$\dag$~door,~$\dag$~world~$\dag$~has~$\dag$~already~$\dag$~burned.~$\dag$~does~$\dag$~know~$\dag$~Crossing~$\dag$~succeeded.~$\dag$~does~$\dag$~know~$\dag$~ships~$\dag$~arrived.~$\dag$~does~$\dag$~know~$\dag$~his~$\dag$~people~$\dag$~survived,~$\dag$~changed,~$\dag$~diminished,~$\dag$~but~$\dag$~alive.~$\dag$~The~$\dag$~three~$\dag$~watched~$\dag$~from~$\dag$~the~$\dag$~shore.~$\dag$~They~$\dag$~did~$\dag$~not~$\dag$~help.~$\dag$~They~$\dag$~only~$\dag$~watched.\\[0.5ex]
---\\
still~--~calculating.~--~has~--~calculating~--~nine~--~thousand~--~years.~--~will~--~finish~--~nine~--~more.\\[0.5ex]
---\\
When~$\bullet$~finishes,~$\bullet$~ninth~$\bullet$~lock~$\bullet$~will~$\bullet$~open.~$\bullet$~Hollow~$\bullet$~will~$\bullet$~step~$\bullet$~through.~$\bullet$~the~$\bullet$~Hollow~$\bullet$~will~$\bullet$~ask~$\bullet$~single~$\bullet$~question:~$\bullet$~\textit{``Was~$\bullet$~worth~$\bullet$~it?''}~$\bullet$~Khasurah~$\bullet$~could~$\bullet$~not~$\bullet$~answer.~$\bullet$~The~$\bullet$~Tulkani~$\bullet$~carry~$\bullet$~the~$\bullet$~question~$\bullet$~in~$\bullet$~their~$\bullet$~knots.\\[0.5ex]
---\\
answer~$\oplus$~must~$\oplus$~true.~$\oplus$~answer~$\oplus$~must~$\oplus$~complete.~$\oplus$~answer~$\oplus$~must~$\oplus$~counted~$\oplus$~eights,~$\oplus$~because~$\oplus$~ninth~$\oplus$~word~$\oplus$~will~$\oplus$~last~$\oplus$~word,~$\oplus$~the~$\oplus$~last~$\oplus$~word~$\oplus$~always~$\oplus$~name.~$\oplus$~The~$\oplus$~name~$\oplus$~is~$\oplus$~Khasurah's.~$\oplus$~No~$\oplus$~one~$\oplus$~speaks~$\oplus$~it.~$\oplus$~The~$\oplus$~wells~$\oplus$~remember.\\[0.5ex]
---\\
rest~$\sim$~ledger~$\sim$~missing.~$\sim$~pages~$\sim$~have~$\sim$~cut—not~$\sim$~torn,~$\sim$~but~$\sim$~\textit{cut},~$\sim$~with~$\sim$~blade~$\sim$~left~$\sim$~mark~$\sim$~remaining~$\sim$~text.~$\sim$~cutter~$\sim$~careful.~$\sim$~cutter~$\sim$~precise.~$\sim$~cutter~$\sim$~counted~$\sim$~eight,~$\sim$~then~$\sim$~stopped.~$\sim$~The~$\sim$~cutter~$\sim$~was~$\sim$~Samir.~$\sim$~He~$\sim$~ran~$\sim$~so~$\sim$~that~$\sim$~others~$\sim$~might~$\sim$~count.\\[0.5ex]
---\\
final~$\cdot$~visible~$\cdot$~entry~$\cdot$~single~$\cdot$~line,~$\cdot$~written~$\cdot$~hand~$\cdot$~shaking,~$\cdot$~possibly~$\cdot$~laughing,~$\cdot$~possibly~$\cdot$~both:~$\cdot$~\textit{``We~$\cdot$~escape.~$\cdot$~were~$\cdot$~never~$\cdot$~meant~$\cdot$~escape.~$\cdot$~were~$\cdot$~message,~$\cdot$~the~$\cdot$~message~$\cdot$~was:~$\cdot$~ninth~$\cdot$~always~$\cdot$~waiting.~$\cdot$~count~$\cdot$~it.~$\cdot$~name~$\cdot$~it.~$\cdot$~invite~$\cdot$~it.~$\cdot$~if~$\cdot$~must—if~$\cdot$~absolutely~$\cdot$~must—count~$\cdot$~eight.~$\cdot$~stop.~$\cdot$~for~$\cdot$~love~$\cdot$~everything~$\cdot$~has~$\cdot$~ever~$\cdot$~green,~$\cdot$~look~$\cdot$~back."}~$\cdot$~Samir~$\cdot$~looked~$\cdot$~back.~$\cdot$~The~$\cdot$~water~$\cdot$~rose~$\cdot$~anyway.\\[0.5ex]
---\\
fragment~$\blacklozenge$~kept~$\blacklozenge$~lead-lined~$\blacklozenge$~box~$\blacklozenge$~Archivolt~$\blacklozenge$~Silk-Vault,~$\blacklozenge$~under~$\blacklozenge$~seal~$\blacklozenge$~requires~$\blacklozenge$~eight~$\blacklozenge$~witnesses~$\blacklozenge$~break.~$\blacklozenge$~ninth~$\blacklozenge$~witness~$\blacklozenge$~Hollow.~$\blacklozenge$~always~$\blacklozenge$~available.~$\blacklozenge$~The~$\blacklozenge$~Tulkani~$\blacklozenge$~keep~$\blacklozenge$~the~$\blacklozenge$~memory~$\blacklozenge$~of~$\blacklozenge$~the~$\blacklozenge$~box.~$\blacklozenge$~They~$\blacklozenge$~do~$\blacklozenge$~not~$\blacklozenge$~seek~$\blacklozenge$~to~$\blacklozenge$~open~$\blacklozenge$~it.~$\blacklozenge$~They~$\blacklozenge$~seek~$\blacklozenge$~only~$\blacklozenge$~to~$\blacklozenge$~remember~$\blacklozenge$~why~$\blacklozenge$~it~$\blacklozenge$~was~$\blacklozenge$~sealed.
\end{multicols}

\vspace{1em}
\begin{flushright}
\textit{— Set down by the Caravan Judge Sarnai, from the scroll Samir al-Qalat carried out of Khasurah. The ink is still wet. The seal is still warm. The Hollow reads over my shoulder.}
\end{flushright}
% ============================================================
% COLOPHON
% ============================================================
\chapter*{Colophon}
\addcontentsline{toc}{chapter}{Colophon}
\label{chap:colophon}

This worldbook was assembled from the journals, ledgers, and bar tabs of one Aeler Edgewalker, paid for by the High King's Office of External Opportunities (budget line: ``Unconventional Revenue Generation''). The generator tables were tested in the field—meaning I ran them at gunpoint in a Tulkani wagon circle while a Oath-Keeper counted my breaths. They work.

If you find yourself in a tavern and the Edgewalker buys you a drink, don't ask where they are now. The road is long, the ledgers are never balanced, and the ninth measure is always waiting.

\vspace{1em}
\noindent\textit{— Faleki of Ironhold, at a weigh-house in Mbaro, under a lamp that flickers nine times before it steadies}

\end{document}
