% !TeX program = lualatex
\documentclass[12pt,oneside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tcolorbox}
\tcbuselibrary{skins, breakable}
\usepackage{makeidx}
\makeindex

% ----- Card suit and rank formatting -----
\renewcommand{\spadesuit}{♠}
\renewcommand{\heartsuit}{♥}
\renewcommand{\clubsuit}{♣}
\renewcommand{\diamondsuit}{♦}

\newcommand{\cardSuit}[1]{%
  \ifx#1\spadesuit \textcolor{black}{\spadesuit}%
  \else\ifx#1\heartsuit \textcolor{red}{\heartsuit}%
  \else\ifx#1\diamondsuit \textcolor{red}{\diamondsuit}%
  \else\ifx#1\clubsuit \textcolor{black}{\clubsuit}%
  \else #1\fi\fi\fi\fi
}

\newcommand{\card}[2]{%
  \tcbox[colback=white, colframe=gray!50, sharp corners, boxsep=1pt, left=2pt, right=2pt, top=1pt, bottom=1pt]%
    {\cardSuit{#1}\ \textbf{#2}}%
}

\newcommand{\spadecard}[1]{\card{\spadesuit}{#1}}
\newcommand{\heartcard}[1]{\card{\heartsuit}{#1}}
\newcommand{\clubcard}[1]{\card{\clubsuit}{#1}}
\newcommand{\diamondcard}[1]{\card{\diamondsuit}{#1}}

% ----- Almanac environment for regional flavor -----
\newenvironment{almanac}
  {\begin{quote}\small\itshape}
  {\end{quote}}

% ----- Cross-reference convenience -----
\newcommand{\secref}[1]{§\ref{#1}}
\newcommand{\chapref}[1]{Chapter~\ref{#1}}

\begin{document}

% ======================================================================
% FRONT MATTER
% ======================================================================
\frontmatter

\title{\Huge Fate's Edge\\[0.3\baselineskip]
\large The Amaranthine Core}
\author{Three Voices of the Road \\ \small (with apologies to the sea)}
\date{\today}
\maketitle

\tableofcontents

\newpage

\chapter*{Preface: Three Voices}
\addcontentsline{toc}{chapter}{Preface}

\noindent
This book is not a travel guide. It is a collection of thresholds: bridges that remember every toll, fog-bells that count your breaths, stones that warm when you speak a true name. Three travelers have walked these roads before you. They have agreed to lend their voices to the pages that follow. Listen to each; they will disagree, and that is the point.

\vspace{1em}
\noindent\textbf{Lady Isolde of the Scar, Fifth of Her Name}
\newline
\textit{``The sun shattered at Lence. The pieces are still falling. Which shadow will catch you?''}

I am Isolde, and I have worn a gorget to three tournaments, buried two brothers, and learned that chivalry is not a code -- it is a weight. You carry it or it crushes you. There is no third option. My family has held the valley north of Rouen since before the Fenwoods burned Lence. In the high season we joust; in the low season we scheme. The Belworth River is our border and our wound. I speak for Vhasia, for Viterra, for the hedge-law courts and the fractured sun. When I tell you a bridge will hold, I have counted its stones. When I warn you a lord cannot be trusted, I have seen the crack in his seal.

\vspace{1em}
\noindent\textbf{The Bravo of Silkstrand (no other name, no other need)}
\newline
\textit{``Every bridge is a promise. Every canal carries a secret. But the best secrets are the ones you share over wine.''}

I am nobody -- or I was, once. The canals of Silkstrand do not care about your bloodline. They care about your stroke with an oar and your aim with a dagger. I started as a dock-rat, became a duelist, and now I am something else: a witness to the city that never sleeps. The Exchange closes at midnight, but the bargains never end. I have won purses and lost friends, and I still smile. Why? Because the sun rises over the laguna every morning, and there is always another bridge to cross. I speak for Silkstrand, for the Brass Coast, for the Sidhi freedmen who taught me that freedom is a story you tell yourself until it becomes true.

\vspace{1em}
\noindent\textbf{The Edgewalker (Aeler, name forgotten by design)}
\newline
\textit{``The mountain breathes. The stone remembers. I am the crack in both.''}

I am an Aeler Edgewalker. You call us mountain-dwellers who measure breath in silver bells and seal our agreements in keystone. How quaint. We don't just measure breath -- we know the load each lung can bear. When a Tulkani chieftain swears an oath on his firstborn? That's not honor -- it's a keystone I've already wedged into three different seals. The first crack was his. The second was mine. When a Veiled Sister binds a demon? That's not terror -- it's a fault line I've traced in the bedrock. The first tremor was hers. The second is coming. You see the surface as ``full of problems.'' I see the weight in the stone, the strain in the beam. Problems are not obstacles -- they are the pressure that shapes the mountain. I speak for the Aeler, for the Mistlands, for the deep vaults where the Hollow is not a myth but a flaw in the foundation. When I tell you a tunnel is sound, I have walked its stress lines. When I tell you to flee, trust the stone's memory.

\vspace{1em}
\noindent
The three of us have walked these roads. We have paid our tolls -- in blood, in silver, in silence. Use the cards to find your own. When you are lost, ask: \emph{What does this place witness? And what does it ask in return?}

Then listen. The road is patient. The Hollow is patient. We are not. But we have left the lamp lit for you.

% ======================================================================
% MAIN MATTER
% ======================================================================
\mainmatter

% ----------------------------------------------------------------------
% PART 0: REFERENCE AND NAVIGATION (LEAN)
% ----------------------------------------------------------------------
\part{Reference and Navigation}

\input{copyright}
\chapter{Reference and Navigation}
\label{chap:reference-navigation}

\section*{How to Use This Chapter}

This chapter provides a structured index and navigation guide to the regions, mechanics, and cartomancy tools of the Amaranthine Core.

\medskip
\noindent
For the cartomancy generator procedure, see \secref{chap:cartomancer-compass}. For travel procedures and the Wilds hexcrawl, see \secref{chap:wilderness}.

\section{Quick Reference by Region}

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Region} & \textbf{Subgenre / Key Theme} \\
\hline
Kahfagia & Creole thalassocracy, signal-lane politics, Lethai-ar infiltration \\
Ecktoria & Imperial decay, bureaucratic horror, the Vigil's three faces \\
Acasia & Mercenary realism, cursed borderlands, the Curse Awakening \\
Vhasia & Courtly intrigue, fractured legitimacy, the wound of Lence \\
Viterra & Legal warfare, frontier justice, the Queen's fragile peace \\
Thepyrgos & Vertical city, synod edicts, war between Witness and Ninth \\
Aeler & Survival horror underground, breath-counting, the pressure that thinks \\
Mistlands & Folk horror, bell-lines, salt as law, Direwood encroachment \\
Ykrul & Nomadic epic, Kon'reh geometry, hostage strings, wolf standards \\
Vilikari & Legal dualism, hybrid identity, frontier pragmatism \\
Black Banners & Mercenary realism, logistics-driven violence \\
Linn & Norse coastal saga, frost-bound oaths, Temple crusade \\
Ubral & Highland clan honor, wergild, geometry of grudges \\
Silkstrand & Mercantile intrigue, layered loyalties, the Ragman's threads \\
The Wilds & Hexcrawl wanderlust, weather as character \\
Dungeons & Living infrastructure, survival horror, forgotten machinery \\
The Ways Between & Dream-death metaphysical, virtue and vice, the Thirteenth Stone \\
\hline
\end{tabular}
\end{center}

\section{Mechanics Index}

\begin{itemize}
  \item \textbf{Cultural Mechanics:} See the ``Cultural Mechanics'' section in each region chapter
  \item \textbf{SB Spend Tables:} See the ``SB Spend Table'' section in each region chapter
  \item \textbf{Boss Seeds:} See the ``Boss Seeds'' section in each region chapter
  \item \textbf{Travel:} \secref{chap:wilderness} for overland; \secref{chap:kahfagia} and \secref{chap:linn} for sea travel
  \item \textbf{Threshold Crossings:} \secref{chap:ways-between}
  \item \textbf{Underground:} \secref{chap:aeler} and \secref{chap:dungeons}
\end{itemize}

\section{Cross-Region Hooks}

\begin{itemize}
  \item \textbf{Acasia--Ecktoria:} The Curse vs. the Flame; imperial ghosts
  \item \textbf{Vhasia--Viterra:} The Belworth border; the Fenwood wound
  \item \textbf{Aeler--Mistlands:} The Protectorate; bell-lines; salt and iron
  \item \textbf{Ykrul--Vilikari:} The Foedus Stone; hostage strings; grazing rights
  \item \textbf{Kahfagia--Silkstrand:} Sea lanes; the Spider's web; dyewater trade
  \item \textbf{Thepyrgos--Viterra:} Synod law vs. hedge-law; the Black River border
\end{itemize}

% ======================================================================
% PART I: THE CARTOMANCER'S COMPASS (FULL)
% ======================================================================

\chapter*{The Cartomancer's Compass}
\label{chap:cartomancer-compass}
\addcontentsline{toc}{chapter}{The Cartomancer's Compass}

\begin{quote}
\textit{``The world does not speak in words. It speaks in suits, ranks, and the spaces between. Every bridge, every bribe, every broken oath can be laid out on a table of felt and dealt from a worn deck. This is not superstition. It is method.''}\\
\hfill --- Mira of the Wandering Circle, Tulkani cartomancer
\end{quote}

\section*{A Tulkani Perspective on the Art}

I was seven when my grandmother first placed a deck in my hands. Not a new deck---a worn one, its edges soft as silk, its faces faded by a hundred thousand shuffles. ``This is your first threshold,'' she said. ``Learn to read it, and you will never be lost.''

The Tulkani do not build temples. We build fires. We do not carve scriptures into stone; we let the cards tell us what the road has seen. Cartomancy is our witch-work, our way of reading the Echo (what has been), steadying the Veil (what is), and leaning into Flow (what wishes to be). It is not fortune-telling. It is witnessing---a conversation with the world's memory, conducted through the language of suits and ranks.

The other peoples of Fate's Edge have their own names for what we do. The Aeler call us ``thread-readers'' and hire us to audit their ledgers. The Ykrul call us ``bone-speakers'' and consult us before every migration. The Ecktorian Inquisition calls us heretics and burns our decks when they find them. We keep shuffling.

This chapter is my gift to the road. It explains how to read the cards as we do---not as a tool for controlling fate, but as a way of listening to what the world has already decided to tell you.

\section*{The Four Suits: A Universal Grammar}

Every card in every generator carries the same four suit-meanings. They are the alphabet of the world's witness. In the Tulkani tradition, we say the suits are the four breaths of the world:

\subsection*{\spadesuit\ Spades: Place --- The Stage}

Where does the scene happen? \spadesuit\ is the threshold, the landmark, the ground beneath the action. It answers: \textit{Where are we?}

\begin{itemize}
    \item \textbf{High rank (10--A):} A major landmark---a fortress, a capital, a haunted ruin. The place has a name that echoes in local songs.
    \item \textbf{Mid rank (6--9):} A functional location---a ford, a market, a shrine. The place has a purpose and a keeper.
    \item \textbf{Low rank (2--5):} A liminal or neglected place---a crossroads, a collapsed bridge, a smuggler's cove. The place is defined by what it lacks.
\end{itemize}

In the regional chapters, \spadesuit\ entries are always places. See, for example, the bridges and canals of Silkstrand or the cairns and passes of Ubral.

\subsection*{\heartsuit\ Hearts: Actor --- The Face}

Who drives the scene? \heartsuit\ is the person, the faction, the voice that speaks. It answers: \textit{Who is here?}

\begin{itemize}
    \item \textbf{High rank (J--A):} An elite, a ruler, a power-broker. The actor has authority and reach beyond the immediate scene.
    \item \textbf{Mid rank (6--10):} A professional, a functionary, a skilled operator. The actor has a trade and a price.
    \item \textbf{Low rank (2--5):} A commoner, an outcast, a victim. The actor needs something or has been failed by someone.
\end{itemize}

In the regional chapters, \heartsuit\ entries are always people or factions. See the merchants and spies of the Brass Coast or the spirit-wardens of the Mistlands.

\subsection*{\clubsuit\ Clubs: Pressure --- The Complication}

What pushes back? \clubsuit\ is the storm, the burden, the knife in the dark. It answers: \textit{What goes wrong?}

\begin{itemize}
    \item \textbf{High rank (10--A):} A catastrophe, a betrayal, an invasion. The complication changes the shape of the journey.
    \item \textbf{Mid rank (6--9):} A serious obstacle---a feud, a quarantine, a broken bridge. The complication costs time, resources, or reputation.
    \item \textbf{Low rank (2--5):} A minor hazard---a sudden squall, a lost token, a raised toll. The complication is a nuisance, not a disaster.
\end{itemize}

In the regional chapters, \clubsuit\ entries are always complications or threats. See the river floods and witness-cord cuts of Ayokha or the market riots of Silkstrand.

\subsection*{\diamondsuit\ Diamonds: Leverage --- The Reward}

What opens the door? \diamondsuit\ is the token, the writ, the secret that changes Position. It answers: \textit{What do you gain?}

\begin{itemize}
    \item \textbf{High rank (J--A):} A major asset---a noble title, a keystone charter, a general pardon. The reward reshapes the party's standing.
    \item \textbf{Mid rank (6--10):} A significant tool---a pass-stone, a witness cord, a lighthouse charter. The reward solves a specific class of problem.
    \item \textbf{Low rank (2--5):} A minor advantage---a bridge token, a guest-cup, a rumor. The reward grants a single use of improved Position.
\end{itemize}

In the regional chapters, \diamondsuit\ entries are always rewards or leverage. See the keystone charters of Ngomebe or the water-share tablets of Fhara.

\section*{Cartomancy as Witchcraft}

In the Tulkani tradition, cartomancy is not separate from the craft---it \textit{is} the craft. Every reading is a Quick Working, performed at a threshold (the table, the crossroads, the moment between decisions). The cards are the tools; the reader is the witness.

\subsection*{The Three Layers of Reading}

Like all witchcraft, cartomancy operates across the three Concordant Layers:

\begin{itemize}
    \item \textbf{Echo (What Has Been):} The cards reveal the accumulated weight of past decisions. A \spadesuit\ from Echo shows the history of a place; a \heartsuit\ from Echo shows a debt that has not yet been paid.
    \item \textbf{Veil (What Is):} The cards show the present moment, the boundary between potential and actual. A \clubsuit\ from Veil is a current threat; a \diamondsuit\ from Veil is an opportunity the party is ignoring.
    \item \textbf{Flow (What Wishes to Be):} The cards reveal the direction of change, the shape of what is coming. A high card from Flow is a promise; a low card is a warning.
\end{itemize}

\subsection*{The Reader's Price}

Every reading exacts a price. The Tulkani say: \textit{``The cards do not give; they lend. And they always collect.''}

When you read the cards, choose one of the Three Prices:

\begin{itemize}
    \item \textbf{Shadow:} You carry the weight of what you have seen. The GM gains 1 SB, which they may spend on a complication tied to the reading.
    \item \textbf{Shame:} The reading reveals something you did not wish to know. You gain a Condition (Guilty, Haunted, or Doubtful) until you act on what you have learned.
    \item \textbf{Identity:} The reading changes you. Mark 1 Identity Strain. You are not quite the same person who sat down at the table.
\end{itemize}

\subsection*{The Cartomancer's Gift}

Once per session, a cartomancer may perform a full reading using the procedure in Section~\ref{sec:generator-procedure}. This is a Full Ritual---it requires Significant Time, a witness, and the willingness to pay the price. On a success, the GM must answer one question about the scene, the journey, or the campaign truthfully. On a partial, the answer is cryptic. On a miss, the cards show a false path---the GM gains 2 SB and may introduce a complication.

\section*{Rank and Severity: The Weight of the Card}

The rank of a card---from Ace (high) to 2 (low)---sets the scale of its suit's meaning. In the Tulkani tradition, we say: \textit{``The rank is the breath; the suit is the voice. A high breath speaks louder; a low breath is a whisper.''}

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Rank} & \textbf{Narrative Weight} \\
\hline
Ace & Legendary, campaign-altering, mythic \\
King/Queen/Jack & Major, faction-level, lasting consequences \\
10--7 & Significant, scene-defining, costly \\
6--4 & Moderate, manageable, a real cost \\
3--2 & Minor, a complication but not a crisis \\
\hline
\end{tabular}
\end{center}

\noindent
\textbf{Timer Conversion:}
\begin{itemize}
    \item 2--5: 4-segment timer (short, local problem)
    \item 6--10: 6-segment timer (standard, journey-length)
    \item J, Q, K: 8-segment timer (extended, regional problem)
    \item A: 10-segment timer (epic, campaign-length)
\end{itemize}

\section*{Card Faces: The Court as Character}

Face cards (Jack, Queen, King) are not merely high ranks. They are \textit{persons} with agency, personality, and often a name. The Tulkani say: \textit{``The court does not sit in judgment of you. You sit in judgment of them.''}

\begin{itemize}
    \item \textbf{Jack:} An agent, a functionary, a messenger. The Jack acts on someone else's authority. They are competent but replaceable.
    \item \textbf{Queen:} A power in their own right. The Queen has influence, resources, and a network. They are not easily replaced.
    \item \textbf{King:} The ultimate authority in their domain. The King is the source of law, the peak of the hierarchy. To challenge a King is to challenge the order itself.
\end{itemize}

When a face card appears in a \heartsuit\ (actor) draw, give the NPC a name, a motive, and a quirk. When it appears in a \spadesuit\ (place) or \clubsuit\ (pressure) draw, treat it as a \textit{personified} force: a King of Clubs might be a feud that has lasted generations; a Queen of Spades might be a bridge that remembers every toll.

\section*{Aces and the Ninth: The Hollow's Attention}

An Ace is not merely the highest rank. It is a \textit{threshold}. In the Tulkani tradition, we say: \textit{``The Ace is the door that opens when you are not looking. Do not step through it lightly.''}

When an Ace appears in any draw, the Hollow takes notice. The GM gains 1 extra Story Beat (beyond any generated by 1s) to spend on a supernatural complication: a ghostly procession, a bell that rings without a clapper, a shadow that walks independently.

In regions where the Ninth Taboo is strong (Theona, Aelinnel, Ways Between), an Ace may also trigger a specific regional omen. For example:

\begin{itemize}
    \item In Theona, an Ace of \spadesuit\ might add a ``missing step'' to a bridge.
    \item In the Ways Between, an Ace of \heartsuit\ might cause a reflection to speak.
    \item In the Mistlands, an Ace of \clubsuit\ might cause a bell-line to fail.
\end{itemize}

The GM should always read an Ace as a signal to escalate the uncanny.

\section*{By the Numbers: Procedure for Any Generator}
\label{sec:generator-procedure}

When you sit down to generate a scene, a journey leg, or a settlement, follow this procedure. The Tulkani say: \textit{``The cards are not a script. They are a conversation. Listen to what they tell you, but do not let them speak for you.''}

\begin{enumerate}
    \item \textbf{Name the destination.} Which region? Which route? Which authority controls the gate?
    \item \textbf{Shuffle and draw.} Draw cards face-up, one for each suit you need. Most generators require all four suits; some (like travel legs) may use only three.
    \item \textbf{Read the \spadesuit.} This is your place. Describe it first. Set the stage.
    \item \textbf{Read the \heartsuit.} This is your actor. Introduce them. What do they want? Why are they here?
    \item \textbf{Read the \clubsuit.} This is your pressure. Complicate the scene. What just went wrong, or what is about to?
    \item \textbf{Read the \diamondsuit.} This is your leverage. What can the players gain if they succeed? Be specific.
    \item \textbf{Set the timer.} Use the highest rank among the four cards to determine the size of the scene's primary timer (see above).
    \item \textbf{Interpret faces and Aces.} Give names to face cards. Apply regional Ace effects.
    \item \textbf{Play to find out.} Use the generated elements as prompts, not scripts. The cards are seeds, not fences.
\end{enumerate}

\section*{Regional Flavors: How Suits Shift Meaning}

While the core suit meanings remain constant, each region inflects them with its own texture. A \spadesuit\ in Ecktoria is a forum or a coin-house; a \spadesuit\ in the Valewood is a phasing ruin or a star-road. A \heartsuit\ in the Black Banners is a mercenary captain or a condotta broker; a \heartsuit\ in Ykrul is a khatun or a bone-singer.

The Tulkani say: \textit{``The same card speaks differently in every land. The fool who reads it the same way twice is a fool twice over.''}

Use the regional chapters as your lexicon. When you draw a card, consult the relevant table. The tables are designed to evoke the specific flavor of the region.

\subsection*{Example: A Draw in Acasia}

You are running a journey through the Broken Marches. You draw:

\begin{center}
\spadesuit 7, \quad \heartsuit J, \quad \clubsuit 4, \quad \diamondsuit Q
\end{center}

\begin{enumerate}
    \item \spadesuit 7: Hill-Motte---Fresh palisade; ditch scattered with caltrops.
    \item \heartsuit J: Bride with No Dowry but a Claim---Daggers in the wedding chest.
    \item \clubsuit 4: Wedding Turns Ambush---You are caught between two bloodlines.
    \item \diamondsuit Q: Marriage Proxy---From the Margravine; bind a hill-king to your cause.
    \item \textbf{Highest rank:} Queen (equivalent to 8). The scene's primary timer is 8 segments: \textit{The Feud Escalates [8]}.
    \item \textbf{Interpretation:} The party arrives at a hill-fort on the eve of a wedding. The bride is caught between two feuding bloodlines. The wedding turns to violence. If the party can survive the ambush and broker peace, the Margravine will grant them a Marriage Proxy---a powerful tool to bind a hill-king to their cause. The 8-segment timer tracks how quickly the feud spirals into open war.
\end{enumerate}

This is the generator at work: four cards, one scene, immediate fiction.

\section*{The Unwritten Suit: When a Draw Is Incomplete}

Sometimes, a generator does not require all four suits. A travel leg may only need \spadesuit, \heartsuit, \clubsuit. A social encounter might only need \heartsuit\ and \clubsuit. When a suit is missing, treat its absence as a question:

\begin{itemize}
    \item \textbf{No \spadesuit:} The place is ambiguous, shifting, or hidden. The scene occurs in a liminal space---a crossroads, a fog-bank, a threshold.
    \item \textbf{No \heartsuit:} The actor is absent, silent, or dead. The pressure and place act without a face. This can be more frightening than a named NPC.
    \item \textbf{No \clubsuit:} The scene is calm, peaceful, or mercifully straightforward. This is rare in Fate's Edge. Treasure it.
    \item \textbf{No \diamondsuit:} The reward is not a token but an outcome. The players gain Position, information, or a cleared timer rather than a physical asset.
\end{itemize}

\section*{Interpreting Any Draw, Any Region}

When you need a prompt in a hurry---a random encounter, a rumor at a tavern, a discovery in a ruin---draw a single card. Use this quick-reference:

\begin{center}
\begin{tabular}{|l|l|}
\hline
\spadesuit & A place or landmark. What is its name? What happened here? \\
\heartsuit & A person or faction. What do they want? What do they fear? \\
\clubsuit & A complication or threat. Who or what is the obstacle? \\
\diamondsuit & A reward, token, or secret. What can be gained? \\
\hline
\end{tabular}
\end{center}

The rank tells you the scale. A low rank is local and personal; a high rank is regional and political. A face card demands a name; an Ace demands the Hollow's attention.

\section*{Tulkani Cartomancy Traditions}

The Tulkani have developed specific practices around cartomancy that differ from the regional generators. These are the traditions my grandmother taught me:

\subsection*{The Traveler's Spread (Three Cards)}

For a journey, the Tulkani lay three cards:
\begin{itemize}
    \item \textbf{First Card:} The road behind. What have you left behind, and what follows you?
    \item \textbf{Second Card:} The road ahead. What waits for you, and what will you face?
    \item \textbf{Third Card:} The road beneath. What is true, whether you wish it or not?
\end{itemize}

This spread is a Quick Working; it requires no roll, but the reader must accept the Shadow price.

\subsection*{The Hollow's Spread (Five Cards)}

For matters of debt, obligation, or the Hollow's attention, the Tulkani lay five cards:
\begin{itemize}
    \item \textbf{First Card:} The weight you carry.
    \item \textbf{Second Card:} The weight you owe.
    \item \textbf{Third Card:} The weight you have forgotten.
    \item \textbf{Fourth Card:} The weight that carries you.
    \item \textbf{Fifth Card:} The weight that will collect.
\end{itemize}

This spread is a Full Ritual. It requires Significant Time, a witness, and the acceptance of Identity Strain. On a success, the GM must answer one question about the nature of the debt. On a partial, the answer is cryptic. On a miss, the reader becomes the debt's next collector---a Complication Timer begins.

\subsection*{The Cartomancer's Deck as a Symbol}

For Tulkani characters, a cartomancer's deck is a Symbol in the Invoker sense. It is a Minor Asset that grants:
\begin{itemize}
    \item Access to the cartomancy rituals described above.
    \item A Patron's Gift (Imbuement) effect: once per scene, the cartomancer may spend an action to draw a card and gain +1 die to a related roll. The GM gains 1 SB.
    \item Obligation to the Hollow: Each reading marks 1 Obligation, as the cartomancer borrows from the world's memory.
\end{itemize}

\section*{Final Guidance: The Cards Are Oracles, Not Scripts}

The generators in this book are tools, not tyrants. If a draw gives you an entry that contradicts established fiction, or if you simply prefer a different interpretation, overrule it. The deck is a prompt, not a prison.

But trust it, too. The most memorable sessions often come from the card you did not expect. The Ace that arrives at the wrong moment. The Jack who speaks a name no one wanted to hear. The diamond that glows with a secret.

The Tulkani say: \textit{``The cards do not lie. But they do not tell the whole truth. The rest is yours to live.''}

Shuffle. Draw. Interpret. And let the road unfold.

\vspace{1em}
\begin{quote}
\textit{``The road is never closed. It is only passed to the next walker.''}\\
\hfill --- Mira of the Wandering Circle, on the eve of a journey she did not survive
\end{quote}

\section*{GM Tools: Using Cartomancy at the Table}

\subsection*{Reading the Cards as a GM}

The cartomancy framework is designed to be used by GMs, not just players. When you need to generate a scene, a complication, or a reward, draw a card and interpret it through the lens of the region and the current fiction.

\textbf{Example:} The party is in Vhasia, negotiating with a border patrol. You draw a \clubsuit 9. The entry says: \textit{``Feud Spark---A name spoken wrong reopens an old debt; blood is the only coin that pays.''} You interpret this as: the patrol captain's family has a blood feud with one of the party's allies. He recognizes the name and demands a price---either a ransom, a duel, or a favor.

\subsection*{Players as Cartomancers}

If a player wants to play a Tulkani cartomancer, use the following framework:

\begin{itemize}
    \item \textbf{Affinity:} Cartomancer's Eye---+1 die to Investigation or Insight rolls involving patterns, hidden connections, or the weight of obligation.
    \item \textbf{Starting Talents:} Cartomancy (4 XP)---grants access to the Traveler's Spread and the Hollow's Spread.
    \item \textbf{Patron:} The Hollow---the cartomancer's Obligation is to the memory of the world itself. The Hollow collects through readings.
\end{itemize}

\subsection*{The Cost of Cartomancy}

Remember: every reading is a transaction. The cards lend, and they collect. A player who reads too often, or who ignores what the cards reveal, will find the Hollow's attention turning their way.

\section*{Colophon}

\textit{This chapter was written in a Tulkani wagon, parked at a crossroads where three roads meet and one does not. The ink is lamp-black and tears. The deck is worn smooth by a hundred thousand shuffles. The ninth card is missing---it always is.}

\textit{---Mira of the Wandering Circle, in the hour before the ninth bell}

\newpage
\chapter{On Obligation and Trade}

The young think economics is about coin. The foolish think it is about resources. The reckless think it is about power. They are all wrong.

Coin is a convenience. Resources are a burden. Power is a promise you have not yet been asked to keep.

The world runs on obligation. Every transaction is a thread. Every trade is a knot. Every gift is a chain that can be pulled---now, or in a generation, or in a century when the one who gave it has been forgotten and the one who received it has become someone else entirely.

We of the Eight Doors understand this. We do not trade. We bind.

\section{The Hollow Economy}

In the lands beyond the settled places, the economy is not measured in silver or steel. It is measured in what is owed. The following are not metaphors. They are the currency of the world.

\subsection{The Debt of Attention}

The most valuable coin is the moment someone looks at you. To be seen is to be recorded. To be recorded is to be remembered. To be remembered is to be owed.

\begin{itemize}
\item \textbf{The Gaze of a Noble}: When a lord looks at you, they have given you something. They have acknowledged your existence. You now owe them. You owe them a return glance, a report, a service, or simply the weight of knowing they saw you.
\item \textbf{The Gaze of a Ghost}: When the dead look at you, the debt is heavier. They have nothing else to give. Their attention is their entire existence. You owe them what they cannot have. You owe them a memory, a story, a name spoken aloud.
\item \textbf{The Gaze of the Ninth}: When the Ninth looks at you, you are no longer a person. You are a character. You are interesting. The debt is infinite. The only payment is a story that ends.
\end{itemize}

\subsection{The Debt of Hospitality}

To eat at someone's table is to borrow their warmth. To sleep under their roof is to borrow their safety. To drink from their cup is to borrow their luck.

Hospitality is not free. It is a loan with no fixed term.

Across the Amaranthine, hospitality is measured in counts:
\begin{itemize}
\item A guest who eats three meals must offer a service in return.
\item A guest who eats nine meals must offer a life-debt.
\item A guest who eats the ninth meal? The Hollow has already counted them.
\end{itemize}

\subsection{The Debt of Gifts}

A gift is not a gift. It is a test. To accept is to agree. To refuse is to declare war.

\begin{itemize}
\item \textbf{The Incomplete Gift}: A gift that is missing something---a blade without a hilt, a cup without a bottom, a coin with one side blank---is a debt in motion. The receiver must complete it. The cost is their own.
\item \textbf{The Shared Gift}: A gift that is passed from one hand to another carries the debts of all who touched it. To accept a shared gift is to accept the obligations of everyone who held it before you.
\item \textbf{The Forbidden Gift}: A gift that cannot be refused---offered by a superior, a patron, or a ghost---is a trap. The debt is immediate. The terms are unknown.
\end{itemize}

\subsection{The Debt of Promise}

A promise is the most dangerous currency. It is pure obligation. It has no physical form. It cannot be seized or destroyed. It can only be kept---or multiplied.

\textbf{The Economics of Promises:}
\begin{itemize}
\item A promise made in public is worth three times a promise made in private.
\item A promise made to a patron is worth five times a promise made to a peer.
\item A promise made to a child is worth ten times a promise made to an adult. The Hollow loves children. They remember everything.
\item A promise made to the dead is worth everything. The dead have nothing else to do but wait.
\end{itemize}

\section{Regional Expressions of Obligation}

Every culture develops its own customs around debt. These customs are not arbitrary; they reflect how that culture interacts with the Ninth and the Hollow. The following table presents a flexible template for designing regional obligation systems. A Game Master can adapt these entries to any setting.

\begin{table}[h]
\centering
\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
\hline
\textbf{Region Type} & \textbf{Primary Currency} & \textbf{Debt Measure} & \textbf{The Ninth's Role} \\
\hline
Agricultural lowlands & Promise & Tally-sticks & The ninth stick is always held by the Hollow \\
Hill enclaves & Hostage string & Blood-tally & The ninth knot is for the Hollow \\
Mountain spires & Cup & Step & The ninth step leads into the Hollow \\
Coastal villages & Bell & Echo & The ninth bell summons the Hollow \\
\hline
\end{tabular}
\caption{Example regional obligation systems.}
\end{table}

\subsection{Adapting the Template}

When designing a region's economic customs, consider three questions:

\begin{enumerate}
\item \textbf{What is the physical token of a debt?} (a stick, a string, a cup, a bell, a coin, a knot, a scar)
\item \textbf{How is a debt measured?} (counts, weight, length, sound, memory)
\item \textbf{How does the Ninth manifest when the debt is ignored?} (a door appears, a string tightens, a cup does not empty, a bell tolls without a ringer)
\end{enumerate}

The answers to these questions will give you a unique economic flavour that is both mechanically useful and narratively rich.

\section{The Real Economy}

Do not let your players think they are trading in coin. The real trade is obligation.

\begin{itemize}
\item \textbf{When a character buys something}: They are not exchanging coin. They are accepting a debt. The debt may be to the seller, to the community, or to the Hollow.
\item \textbf{When a character accepts a gift}: They are accepting an obligation. The gift will be called in. The Hollow will remember.
\item \textbf{When a character offers a service}: They are creating a debt. The debt may be repaid now, or later, or never. Never is not an option.
\end{itemize}

\subsection{The Hollow's Interest}

The Hollow does not charge interest. It charges attention. Every debt that is not paid becomes more visible. The Hollow notices.

\begin{itemize}
\item \textbf{A small debt unpaid for a day}: The Hollow sends a reminder---a knock, a whisper, a sudden chill.
\item \textbf{A medium debt unpaid for a week}: The Hollow sends a witness---a stranger who knows too much, a dream that does not fade.
\item \textbf{A large debt unpaid for a month}: The Hollow sends a collector---a figure who does not leave, a weight that does not lift.
\end{itemize}

\section{Using Debt Timers}

The Ninth and the Hollow work best when they are not random. They are cumulative. Use timers to track the pressure of unfulfilled obligations. The following table offers a general framework that can be adapted to any campaign.

\begin{table}[h]
\centering
\begin{tabular}{|p{3cm}|p{2cm}|p{5cm}|}
\hline
\textbf{Timer} & \textbf{Size} & \textbf{What It Tracks} \\
\hline
Promise & 4--6 & Unkept oaths; ticks when a promise is broken or ignored. When full, the Hollow sends a reminder. \\
Debt & 4--6 & Unpaid obligations of any kind; ticks when a debt is acknowledged but not settled. When full, the Hollow sends a collector. \\
Attention & 6--8 & Escalation; ticks when characters take the ninth step, pour the ninth cup, or ignore warnings. When full, the Ninth manifests physically. \\
\hline
\end{tabular}
\caption{Timer recommendations for tracking obligation.}
\end{table}

\section{GM Guidance: The Balance of Obligation}

The economics of obligation are not punishments. They are consequences. They should never feel arbitrary, and they should never feel unfair. When a character attracts the Ninth or the Hollow through economic entanglement, it should be because they made a choice.

\subsection{The Golden Rules}

\begin{enumerate}
\item \textbf{If a player could not have known what they were doing, do not invoke the Ninth or Hollow.} Let it be a consequence of choice, not of ignorance.
\item \textbf{When the Ninth or Hollow appears, it should change the situation.} A complication, not an ending. The players should still have agency. They should still be able to act. They should simply have to act around the debt.
\item \textbf{Debt is a story, not a number.} Do not reduce it to a ledger. The weight of obligation is felt in whispers, in dreams, in the slow accumulation of dread.
\end{enumerate}

\subsection{Designing Regional Consequences}

When a debt timer fills in a specific region, consider what form the Hollow's attention might take. Use the region's customs to guide you:

\begin{itemize}
\item If the region uses tally-sticks, the Hollow might appear as a figure holding a stick with the character's name.
\item If the region uses hostage strings, the Hollow might tighten an invisible cord around the character's wrist.
\item If the region uses cups, the Hollow might present a cup that must be filled with a story or a memory.
\item If the region uses bells, the Hollow might ring a bell that cannot be unheard.
\end{itemize}

\subsection{A Final Warning}

The economy of obligation is not a system to be solved. It is a texture to be felt. The more you explain it, the less it works. Let the players discover it through experience, through the weight of their own choices.

Do not tell them the cost of a gift. Let them feel it when the gift is called in.

Do not tell them the terms of a promise. Let them learn them when the promise is broken.

The Hollow is patient. It will wait until the debt is interesting.

\vspace{1em}
\hfill --- Matron Alfiya, Keeper of the Eight Doors

\newpage
% ======================================================================
% PART II: THE AMARANTHINE SEA & NORTHERN SHORES
% ======================================================================
\part{The Amaranthine Sea \& Northern Shores}

\begin{quote}
  \small
  \textit{``The sea keeps a better memory than any scribe. The reef remembers every secret whispered above its depths. Do not ask which flag flies highest -- ask which hull leaks slowest.''}
  \\ --- The Bravo of Silkstrand
\end{quote}

% -- Kahfagia --
\chapter{Kahfagia}
\emph{The Flag of Storms.}

\noindent
\textit{``Kahfagia runs on lantern-light and rust, where the lies are as polite as the lanterns that never go out. Every beacon is a promise -- and every reef remembers the ships that broke theirs. Come for the salt fish, stay for the gossip, but never trust a calm tide: the commodores know the lanterns fear the dark, and I've seen a Mirror-Keeper weep when her reflection showed her a daughter she'd sold for a pilot's license.''} \\
--- The Bravo of Silkstrand

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Salt fish porridge -- dried cod soaked overnight, boiled with hardtack and a splash of lamp oil. Every Kahfagian child learns to taste the salt and know which coast their grandmother came from.

\textbf{What does a child learn before they can walk?} To count ships by their lanterns. Three lanterns means a Kahfagian merchant -- safe. Four means an Oshiiran grain barge -- neutral. One, low in the water, means a privateer. Run.

\textbf{What is the first thing you notice about a stranger?} Their hands. Dock-workers have rope-burned palms. Merchants have ink-stained fingers. Privateers have no fingernails -- they bite them off in storms.

\textbf{What is the most common crime?} False manifests. Every third ship carries cargo that doesn't appear on its papers. The Admiralty knows. The Admiralty takes a cut. The Admiralty calls it a ``tariff surcharge.''

\textbf{What does no one talk about but everyone knows?} The Mirror-Keepers are not born -- they are made. And the making requires a death. A sister, a daughter, a lover. Someone who drowned so the keeper could see through the fog.

\textbf{Where do people go when they die?} The tide takes them. Kahfagia has no cemeteries above ground. The drowned are weighted with chain and dropped in the deep channel. Their families light candles in paper boats and float them out at dusk.

\textbf{How do you apologize for a serious wrong?} You offer your bell. Every Kahfagian carries a small brass bell -- the sound of their name. Offering it means ``I give you my voice.'' Refusing to return it is a declaration of blood feud.

\textbf{What is the most important festival?} The Beacon Lighting (midwinter). Every lighthouse keeper lights a new lamp from the Admiral's flame. If your beacon catches first, your year will be lucky. If it sputters, you will owe a death to the sea.

\textbf{What is the secret that could destroy a powerful person?} The Admiral's lantern code has a ninth signal -- never used, never spoken, burned into every keeper's memory. It means \textit{``Kill the witnesses. Sink the evidence. The sea will hide the rest.''}
\end{almanac}

\input{../generators/kahfagia.tex}
\index{Kahfagia}

\newpage

% -- Ecktoria --
\chapter{Ecktoria}
\emph{Marble \& Fire.}

\noindent
\textit{``Ecktoria is a marble coffin for a king who never was. The Everflame burns not for light but for weight -- it measures the lie in every breath. The statues are not watchers -- they are the pressure points in the stone. The undertakers know the body is cold, but the pressure remains. The mountain does not forget what the marble tries to hide.''} \\
--- Aeler Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Spiced flatbread dipped in olive oil and honey -- but only if you can afford the honey. The poor eat millet porridge with a single dried fig. The fig is counted as part of the household's grain allowance.

\textbf{What does a child learn before they can walk?} To count coins by touch. Every Ecktorian child has a box of worn sesterces; by age five, they can tell a debased copper from pure silver by the weight alone.

\textbf{What is the first thing you notice about a stranger?} Their signature. Ecktorians judge by handwriting before face. A clerk with a steady hand is trustworthy. A noble with a flourish is hiding something. A merchant who signs with an X is not worth doing business with.

\textbf{What is the most common crime?} Tax evasion -- but it is not a crime, it is a sport. Every household keeps two ledgers: the white one for the Censor, the red one for the family. The art is in keeping the red ledger from looking like a confession.

\textbf{What does no one talk about but everyone knows?} The Censor's clerks are paid to find discrepancies. They are also paid to lose them. The price of a lost discrepancy is exactly 5\% of the tax owed. This is not corruption. It is ``administrative flexibility.''

\textbf{Where do people go when they die?} The family crypt, beneath the basilica floor. The wealthy are burned, their ashes mixed with marble dust and pressed into the floor tiles. The poor are buried in unmarked graves outside the walls. The Everflame's light does not reach that far.

\textbf{How do you apologize for a serious wrong?} You petition the Censor to have the wrong stricken from the public record. The petition costs a tenth of your annual income. If accepted, the wrong is erased -- legally, it never happened. If rejected, the wrong is published, and you lose your civic standing.

\textbf{What is the most important festival?} The Triumph of the Flame (autumn equinox). The basilica's central flame is carried in procession through every district. To touch the flame is to be absolved of petty debts. To be burned by it is a sign that your debts are not petty.

\textbf{What is the secret that could destroy a powerful person?} The Grand Magistrate's seal is a forgery. The original cracked in 783 AR. The current seal was carved by a prison engraver who was executed to keep the secret. The engraver's apprentice is still alive. They live in Silkstrand. They have the original sketches.
\end{almanac}

\input{../generators/ecktoria.tex}
\index{Ecktoria}

\newpage

% -- Acasia --
\chapter{Acasia}
\emph{The Broken Marches.}

\noindent
\textit{``Acasia was the first province the Utaran Empire swallowed and the last it spat out. Now it is a wound that will not close -- a place where roads loop back on themselves and bridges lead to the same shore. The Curse is not a spell. It is a memory, and the land has a very long memory. In these marches, memory is the only currency that does not devalue.''} \\
--- Lady Isolde of the Scar

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Hardtack soaked in broth -- if you have broth. If not, hardtack and hope. The granaries have been empty for three winters. The merchants sell grain at famine prices. The duke says the roads are too dangerous for caravans. The farmers say the duke owns the caravans.

\textbf{What does a child learn before they can walk?} To name the dead. Every Acasian child has a list of ancestors they repeat at bedtime. The list gets shorter each generation. Not because children forget. Because the dead are forgetting themselves.

\textbf{What is the first thing you notice about a stranger?} Whether they have a shadow. The Curse takes shadows first. A man without a shadow is not cursed -- he is already gone. His body just hasn't stopped moving yet.

\textbf{What is the most common crime?} Bridge-toll evasion. The roads are unsafe, the ferries are unreliable, but the barons still demand payment for crossings no one can make. The crime is not in refusing to pay. The crime is in being caught trying.

\textbf{What does no one talk about but everyone knows?} The Curse began with a promise broken on the Blackwood Bridge. A baron promised his daughter to a lord of the old forest. He married her to a lowland knight instead. The forest remembered. The forest is still waiting for the wedding.

\textbf{Where do people go when they die?} Into the Blackwood. Burial is forbidden -- the ground rejects the dead. Bodies are carried to the tree-line and left. The wood takes them. Sometimes, at dusk, you can see them walking between the trunks. They do not look back.

\textbf{How do you apologize for a serious wrong?} You walk the Broken Mile -- a stretch of road that collapsed into the river a decade ago. If you reach the far side, you are forgiven. If you fall, you were not sorry enough. No one has reached the far side in seven years.

\textbf{What is the most important festival?} The Counting of the Names (midwinter). Every village lists its dead aloud, from oldest to newest. If a name is forgotten, the ghost walks until someone remembers. The list grows longer each year. The memory grows shorter.

\textbf{What is the secret that could destroy a powerful person?} The duke's son is a ghoul. He was lost in the Blackwood as a child. They found him three days later, pale and smiling. He does not age. He does not eat. He only watches. The servants know. The servants do not speak.
\end{almanac}

\input{../generators/acasia.tex}
\index{Acasia}

\newpage

% -- Vhasia --
\chapter{Vhasia}
\emph{The Fractured Sun.}

\noindent
\textit{``The sun shattered at Lence. The pieces are still falling. Which shadow will catch you? The Fenwoods burned the capital a decade ago, but the fire never went out -- it just moved indoors, into every court, every negotiation, every marriage bed. Chivalry is the mask; the knife is the truth.''} \\
--- Lady Isolde of the Scar

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Dark bread and watered wine -- the wine cuts the dust from the road. The nobility eat spiced eggs and fresh cheese, but they eat behind closed doors. To eat well in public is to invite envy. Envy is a dagger.

\textbf{What does a child learn before they can walk?} Which colors not to wear. Black and gold are the Northern Crown's colors. Red and silver are the Southern Parlement's. A child who wears the wrong color to the wrong village may not return home.

\textbf{What is the first thing you notice about a stranger?} Which side of the road they walk on. The north side is the Crown's. The south side is the Parlement's. The middle is no man's land -- literally. Bandits, spies, and messengers use the middle. So do ghosts.

\textbf{What is the most common crime?} Raiding across the Belworth. The river is the border, the border is a suggestion, and the suggestion is enforced by armed men on both sides. The crime is not in raiding. The crime is in being caught on the wrong bank at dawn.

\textbf{What does no one talk about but everyone knows?} The Sun-court was not burned. It was bankrupted. The Fenwoods did not sack Lence -- they bought it. Every judge, every clerk, every guard captain had a price. The Fenwoods paid it. The Vhasians are still paying interest.

\textbf{Where do people go when they die?} To the crypts beneath the Sun-court -- if they can afford a plaque. The poor are buried in unmarked graves along the Belworth floodplain. The river floods every spring. The dead wash downstream, into the land of the other crown. Neither side claims them.

\textbf{How do you apologize for a serious wrong?} You offer a chevauchée -- a raiding party that burns a village in your name. The apology is not in the burning. The apology is in the promise that the village will be rebuilt with your own hands.

\textbf{What is the most important festival?} The Sun-return (spring equinox). Both crowns claim the sun as their symbol. On this day, both hold processions. The processions are identical. The participants do not acknowledge each other. The sun does not care.

\textbf{What is the secret that could destroy a powerful person?} The Queen-Mother has no claim to the regency. Her son is not the Dauphin. The real Dauphin died in childbirth. The child on the throne is a foundling, bought from a Tulkani caravan. The Tulkani know. The Tulkani are still waiting to be paid.
\end{almanac}

\input{../generators/vhasia.tex}
\index{Vhasia}

\newpage

% -- Viterra --
\chapter{Viterra}
\emph{The Hedge-Law Realm.}

\noindent
\textit{``Viterra hums with the quiet hum of old promises kept. The Queen's Peace is not a word -- it is a wax seal warm from the press, and every seal tells a story. The hedge-courts don't keep the peace with swords -- they keep it with second chances. And in Viterra, a second chance is always worth the price of a smile.''} \\
--- The Bravo of Silkstrand

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Barley porridge with a drizzle of honey -- if you own bees. If not, porridge with salt. Salt is the currency of the poor. Honey is the currency of the hopeful.

\textbf{What does a child learn before they can walk?} To recite their lineage back three generations. Every Viterran child knows their grandparents' names, trades, and debts. A child who cannot recite is a child without standing. Standing is everything.

\textbf{What is the first thing you notice about a stranger?} Whether they carry a seal ring. A seal ring means they have the right to make contracts in their own name. A stranger without a seal ring is either a servant or a criminal. In Viterra, the difference is a matter of paperwork.

\textbf{What is the most common crime?} Boundary disputes. A hedge moved three feet east. A milestone replaced under cover of darkness. A ford declared ``private'' when it has been public for two centuries. The courts are clogged. The lawyers are rich. The hedges are patient.

\textbf{What does no one talk about but everyone knows?} The Queen is not a warrior. She has never drawn a sword. Her reputation was forged by a propaganda campaign run by her spymaster. The spymaster is dead. His successor is more discreet.

\textbf{Where do people go when they die?} To the family plot -- if they have a family. The landless poor are buried in parish cemeteries, their headstones marked with a single word: ``Anonymous.'' The ghosts wander the hedgerows at dusk. They do not remember their names. They remember their debts.

\textbf{How do you apologize for a serious wrong?} You petition the hedge-court for a ``writ of contrition.'' You state your wrong aloud while holding a handful of earth from the wronged party's land. If the earth crumbles, you are forgiven. If it holds, you are not.

\textbf{What is the most important festival?} The Opening of the Rolls (autumn). Every county sends its tax records to the capital. The documents are carried by children, so the next generation learns the weight of obligation.

\textbf{What is the secret that could destroy a powerful person?} The Codex of Precedents has a missing page. It was torn out during a trial for treason. The page contained the precedent for executing a monarch. The Queen knows. The Justiciars know. The page is hidden in a lead-lined box beneath the hedge-court at Thornwood.
\end{almanac}

\input{../generators/viterra.tex}
\index{Viterra}

\newpage

% -- Thepyrgos --
\chapter{Thepyrgos}
\emph{City of a Thousand Stairs.}

\noindent
\textit{``Thepyrgos is a vertical labyrinth of stair-rights, rope-guilds, and synod edicts that shift with the tide. The Synod rules with the delicacy of a brick wrapped in velvet -- they hate unsanctioned magic, but they adore infrastructure. Beneath the city, the Unfinished Stair waits for those who ask a question the Synod cannot answer.''} \\
--- The Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Flatbread with olive oil and a boiled egg -- if you are a clerk. If you are a student, bread and water. If you are a professor, bread, water, and a single fig. The fig is a status symbol. The status is ``I have not been audited.''

\textbf{What does a child learn before they can walk?} To read a bell schedule. Thepyrgos runs on bells -- dawn bells, market bells, curfew bells, trial bells. A child who cannot read the bells cannot cross the city without a guardian. A child without a guardian is a child who will be sold to the university for dissection.

\textbf{What is the first thing you notice about a stranger?} Their step count. Thepyrgos has a thousand stairs. A stranger who counts them aloud is either a pilgrim or a spy. A stranger who climbs them in silence is either a native or a ghost.

\textbf{What is the most common crime?} Unsanctioned magic. The Synod licenses every spell cast within the walls. An unlicensed cantrip is a misdemeanour. An unlicensed healing is a felony. An unlicensed summoning is a death sentence -- not for the caster, who will be executed, but for the summoned, who will be bound to the city's foundation as a ``structural consultant.''

\textbf{What does no one talk about but everyone knows?} The Silent Stair is not silent. It whispers the confessions of dead scholars -- their secrets recorded in the stone. The Chain-Lanterns listen. They do not share.

\textbf{Where do people go when they die?} To the catacombs beneath the university. The catacombs are a library of bones. Each skeleton is labeled with the scholar's name, their thesis, and the date of their death. The unlabelled bones are the ones who died before they could finish. They are still writing.

\textbf{How do you apologize for a serious wrong?} You submit a ``retraction'' to the Synod's Committee on Ethical Correction. The retraction is published in the public square, read aloud by a herald, and witnessed by three notaries. The fee is 50 silver. The shame is free.

\textbf{What is the most important festival?} The Convocation of the Steps (summer solstice). Every citizen climbs the thousand stairs to the upper citadel. The Archon reads the names of the dead from the past year. The names are carved into the steps as they are read. The steps are getting crowded.

\textbf{What is the secret that could destroy a powerful person?} The Archon is a fraud. He never passed the Synod's examination. His degree was purchased from a disgraced dean whose signature is forged on every document. The forger is still alive. They live in the catacombs. They have the original quill.
\end{almanac}

\input{../generators/thepyrgos.tex}
\index{Thepyrgos}

% ======================================================================
% PART III: THE NORTHERN REACHES
% ======================================================================
\part{The Northern Reaches}

\begin{quote}
  \small
  \textit{``The mountain counts your breaths. The frost writes your name in ice. In the north, the dead do not wait -- they walk, and they remember every promise you made to keep warm.''}
  \\ --- The Edgewalker
\end{quote}

% -- Aeler --
\chapter{Aeler}
\emph{Crowns \& Under-Vaults.}

\noindent
\textit{``The Aeler hold the mountains by breath, bell, and load-bearing stone. Air is measured, water is bartered, light is rationed -- and every tunnel has a breaking point. I've mapped most of them. The ones I haven't are still settling.''} \\
--- The Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Stone-bread and cave cheese -- the bread is baked on heated stone slabs, the cheese is aged in sealed galleries. The cheese has been aging for three generations. It is an heirloom. It is also barely edible.

\textbf{What does a child learn before they can walk?} To count their breaths. Aeler children wear breath-counters -- small brass bells. Eight breaths, ring the bell. Nine breaths, run. A child who cannot count will not survive the winter.

\textbf{What is the first thing you notice about a stranger?} Their bell. Every Aeler carries a bell tuned to their breath. A stranger's bell rings at a different pitch. A mismatched bell is a warning. A silent bell is a threat.

\textbf{What is the most common crime?} Air theft. In the deep vaults, ventilation is shared. A sealed door, a blocked vent -- these are thefts of breath, punishable by breath-debt. The guilty party donates their own air ration to the victim. The donation is measured in days. The days are counted by bells.

\textbf{What does no one talk about but everyone knows?} The True Masons are not neutral. They are the Vault-Queen's eyes. Every ``independent'' repair, every ``neutral'' inspection -- the True Masons are watching. They do not report to the Queen. They are the Queen.

\textbf{Where do people go when they die?} To the Deep Vaults. The dead are sealed in niches carved into the walls, their names chiselled above their heads. The niches are reused every century. The names are never erased. The dead share walls with strangers. The strangers are patient.

\textbf{How do you apologize for a serious wrong?} You offer a keystone -- a load-bearing stone from your own holding. It is installed in the wronged party's wall. If it holds, you are forgiven. If it cracks, the claim stands. If it crumbles, your holding collapses. Apology is architecture. Architecture is law.

\textbf{What is the most important festival?} The Ringing of the Underways (deep winter). Every bell in every hold is rung eight times. The sound travels through the stone, a heartbeat that reminds the mountain that the living are still here. The dead listen. The dead do not ring back.

\textbf{What is the secret that could destroy a powerful person?} The Vault-Queen is a construct. Assembled after the real queen died in a cave-in. Her heart is a keystone. Her breath is a bellows. The True Masons have the schematics. The schematics are missing three pages -- the pages that describe how to turn her off.
\end{almanac}

\input{../generators/aeler.tex}
\index{Aeler}

\newpage

% -- Mistlands --
\chapter{The Mistlands}
\emph{Bells, Salt, and Breath.}

\noindent
\textit{``The Mistlands lend their name to the Aeler, but the fog is its own master. It hides not just the land, but the law that should bind us. The bell-lines ring to keep the wraiths back, and the Aeler ask for their due -- but when the coin is gone, they take their bells and leave. Then the fog does not forget; it remembers every name they left behind in the damp and dark.''} \\
--- Lady Isolde of the Scar

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Salt fish and sour bread -- the fish is smoked over peat fires, the bread is leavened with fog-damp yeast. The yeast is caught in cloth traps hung from bell-towers. The yeast is not alive. It is not dead either.

\textbf{What does a child learn before they can walk?} To ring a bell. Every Mistlands child is given a small bronze bell at birth. By age three, they can ring the eight notes of the ward-chime. The ninth note is never taught. Discovery is fatal.

\textbf{What is the first thing you notice about a stranger?} Their salt-pouch. A stranger without a pouch is a fool. A stranger with an empty pouch is a fool about to become a wraith.

\textbf{What is the most common crime?} Ward-salt theft. The Aeler control the salt pans, and the salt is the only thing that keeps the dead from rising. A stolen pouch buys a week's protection. The price is a year's labor in the deep galleries. The cycle is perpetual.

\textbf{What does no one talk about but everyone knows?} The Bell-Wardens are not chosen -- they are volunteered. The training holds are in the mountains. The mountains are cold. The ones who return do not speak of what they learned. Their bells ring at a different pitch.

\textbf{Where do people go when they die?} To the fog. Bodies are left at the edge of the bell-line. The mist takes them. Sometimes, at dusk, you can see them walking the lanes. They do not speak. Their presence is a question.

\textbf{How do you apologize for a serious wrong?} You ring the Bell of Atonement -- a cracked bronze bell in every village square. Ring it once for each wrong confessed. If the bell rings true, you are forgiven. If it shatters, you are banished to the fog. The banished never return. Their voices still answer from the white.

\textbf{What is the most important festival?} The Rehanging of the Lines (autumn). Every bell-line is taken down, cleaned, and rehung. The villages are silent for three nights. The dead listen. The dead wait. The dead do not ring.

\textbf{What is the secret that could destroy a powerful person?} The Legate of the Mists is a wraith. She died in 803 AR, during a bell-line failure. The Aeler rebuilt her from fog and memory. She does not know she is dead. She signs documents. Her signature fades after sunset.
\end{almanac}

\input{../generators/mistlands.tex}
\index{Mistlands}

\newpage

% -- Ykrul --
\chapter{Ykrul}
\emph{Wolf Standards, Winter Camps.}

\noindent
\textit{``The Ykrul ride the Violet Steppe, where the sky is a cup and the horizon its lip. Their Kon'reh is not a code but a compass -- it reads the land in steps, in stops, in the weight of a breath. Every crossing has its cost, every ford its toll, every feud its balance in the space between hoof and earth. But the Steppe does not deal in coin; it deals in the weight of a promise and the length of a life.''} \\
--- Aeler Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Fermented mare's milk and dried meat -- the milk is drunk warm from the skin, the meat is sliced thin and dried on saddle-pommels. The meat carries the taste of horse sweat. The milk carries the taste of the steppe.

\textbf{What does a child learn before they can walk?} To sit a horse. Ykrul children are placed on a horse before they can stand. A child who falls learns to hold tighter. A child who cannot ride cannot eat.

\textbf{What is the first thing you notice about a stranger?} Their braids. Red for the Wolf Clan, black for the Iron Horse, gray for the Salt Steppe. A stranger whose braids are cut has been shamed. A stranger with no braids has no clan. A stranger with no clan is prey.

\textbf{What is the most common crime?} Horse theft. A stolen horse is a stolen relative. The price is blood -- the thief's or the owner's. The evidence is the horse's memory. Horses do not lie.

\textbf{What does no one talk about but everyone knows?} The Khagan is not chosen by blood. He is chosen by the horse. Every five years, the clan chiefs lead their best stallions to the sacred ring. The horse that steps forward first chooses the next Khagan. The horse has not stepped forward in twenty years.

\textbf{Where do people go when they die?} To the sky. Bodies are placed on open platforms, left for the wind and birds. The bones are ground into powder, mixed with clay, fired into beads. The beads are worn by the dead's descendants. The beads are warm. The warmth is memory.

\textbf{How do you apologize for a serious wrong?} You offer a hostage string -- a braided cord that ties your wrist to the wronged party's. The cord is cut when the apology is accepted. If the cord is never cut, you are bound. The binding is lifelong. The binding is called ``friend.''

\textbf{What is the most important festival?} The First Grass (spring equinox). The herds are released onto new pasture. The riders race to the horizon and back. The winner names the summer camps. The loser tends the sick animals. The loser always wins next year.

\textbf{What is the secret that could destroy a powerful person?} The Sky-Speaker is a fraud. She cannot call storms. She cannot read omens. She is a horse-witch who learned to read the clouds. The clouds are reliable. The Khagan does not want to know.
\end{almanac}

\input{../generators/ykrul.tex}
\index{Ykrul}

\newpage

% -- Vilikari --
\chapter{Vilikari}
\emph{Laurels \& Longhouses.}

\noindent
\textit{``The Vilikari live in the space between marble and felt, where treaties are not written in law but woven in need. The emperors gave us their chisel, the chieftains gave us their knot -- and where they meet, we find a strength that is not in the stone or the string, but in the space between. We do not forget the argument; we remember that the space between is where we build.''} \\
--- The Bravo of Silkstrand

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Oatmeal with dried berries -- the berries are harvested from the border scrub, where the empire's orchards meet the steppe's wild brambles. The berries taste of both. The oatmeal tastes of neither.

\textbf{What does a child learn before they can walk?} To speak both languages -- Imperial Utaran and the steppe cant. A child who speaks only one will never cross the border. A child who crosses the border comes back with stories. Listening is power.

\textbf{What is the first thing you notice about a stranger?} Which side of the road they ride on. The empire's roads are measured from the center. The steppe's paths are measured from the edge. A stranger who rides in the ditch is a spy.

\textbf{What is the most common crime?} Treaty-splitting -- selling the same rights to both sides. The crime is not in the selling. It is in getting caught before you cross the border.

\textbf{What does no one talk about but everyone knows?} The Foedus Stone is a copy. The original was stolen by a Vilikari trader who owed favors on both sides. She melted it down and sold the marble dust as ``treaty-strengthening powder.'' The treaties are still holding.

\textbf{Where do people go when they die?} To the border. Buried at the edge -- close enough for imperial priests to pray, close enough for steppe shamans to chant. The dead are planted face-up, so they can see both skies. The skies are different colors. The dead do not complain.

\textbf{How do you apologize for a serious wrong?} You invoke the mallus -- the oak-ring court. The wrongdoer circles the ring three times, speaking their apology. On the third circuit, the elders judge. If rejected, the ring closes. Outside is the lawless land. The lawless land is where the wolves are.

\textbf{What is the most important festival?} The Renewal of the Foedus (summer solstice). The stone is washed with milk and honey. The milk is from a Ykrul mare. The honey is from an imperial apiary. The ceremony takes all day.

\textbf{What is the secret that could destroy a powerful person?} The Foedus Stone has a crack. It was repaired with clay and prayer. The clay is drying. The prayer is fading. When the stone splits, the treaty splits with it. No one knows what happens then.
\end{almanac}

\input{../generators/vilikari.tex}
\index{Vilikari}

\newpage

% -- Black Banners --
\chapter{Black Banners}
\emph{Condotta \& Crowns.}

\noindent
\textit{``Mercenary companies trade steel for coin -- today's shield, tomorrow's spear. Ecktorian treasuries do not contain the border war; they stoke it with gold. Between them walk the Ykrul clans, the Vilikari tribes, and the legionaries who remember when a word was a bond.''} \\
--- Aeler Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Hardtack and camp stew -- the stew is whatever was caught, killed, or traded for before dawn. The cook is called ``The Alchemist.'' No one asks their real name.

\textbf{What does a child learn before they can walk?} To count coin. Camp children tally pay shares before they can read. A child who miscounts goes hungry. A child who goes hungry learns to count faster.

\textbf{What is the first thing you notice about a stranger?} Their colors. Every company has its own banner -- lantern, thorn, bell, fishbone, broken scale. A stranger with no colors has no company. A stranger with no company is a deserter. A deserter is a corpse.

\textbf{What is the most common crime?} Desertion -- but desertion is a negotiation. A soldier who wants to leave will sell their gear, their secrets, and their silence. The price is always ``one more season.''

\textbf{What does no one talk about but everyone knows?} The Black Colonel is not a person. The title is held by a wax seal in a lead-lined box. The box is carried by a mute lieutenant. The lieutenant signs documents with the Colonel's thumb-print. The thumb-print is a forgery. The forger is the lieutenant. The lieutenant has no thumbs.

\textbf{Where do people go when they die?} To the pyre. Camp dead are burned in a bonfire of broken spears. The quartermaster charges a fee from the dead's final pay. The final pay is never enough.

\textbf{How do you apologize for a serious wrong?} You stand a watch. The wronged party chooses the hour -- midnight, dawn, the dead of winter. If you survive, you are forgiven. If you do not, the claim is paid.

\textbf{What is the most important festival?} Payday -- but payday is a negotiation. The captain promises payment on the first of the month. The quartermaster promises when the supply train arrives. The soldiers promise loyalty until the payment comes. The payment never comes. The loyalty never ends.

\textbf{What is the secret that could destroy a powerful person?} The Banner-Captain's contract is written in disappearing ink. The ink was supplied by a Sidhi forger paid in promises. The promises were not kept. The forger is now a camp follower. The forger knows where the contract is hidden.
\end{almanac}

\input{../generators/black_banner.tex}
\index{Black Banners}

\newpage

% -- Linns --
\chapter{The Linns}
\emph{Skerries \& Storm-Oaths.}

\noindent
\textit{``North of the Aberderrin Sea, where fjords split the coast and skerries break the tide, dwell the Linn. They are not just traders and raiders; they are keepers of the storm's oath and the sea's weight. Their longships fill the whale-road and the river mouths, not for plunder, but for the promise of the tide. They swear oaths to storm and sea -- not as words, but as a weight that binds, a clause that the deep remembers. They do not pay in silver; they pay in stories, because a tale well-told is the only treasure that lasts when the gold is gone.''} \\
--- Lady Isolde of the Scar

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Seal blubber and flatbread -- the blubber is chewy, the bread is dense, and both are stored in the same barrel. The barrel is shared with the crew. The crew shares everything. Family is obligation.

\textbf{What does a child learn before they can walk?} To tie a knot. Linn children are given a length of rope at birth. By age five, they can tie the eight sea-knots. The ninth knot is for emergencies -- and for funerals. The ninth knot is for when the sea asks for a life.

\textbf{What is the first thing you notice about a stranger?} Their breath. Linn sailors chew kelp to sweeten their breath. A stranger whose breath smells of salt has been at sea. A stranger whose breath smells of copper has spoken a false oath. The sea will answer.

\textbf{What is the most common crime?} Oath-breaking -- but oath-breaking is a storm. The oath-breaker's ship will founder. Their children will be born with webbed fingers. The sea is patient. The sea does not forget.

\textbf{What does no one talk about but everyone knows?} The Sea-Queen is not a queen. She is a figurehead -- literally. Her body is carved from the keel of a ship that sank in 612 AR. Her voice is a recording. The jarls pretend. The pretending keeps the peace.

\textbf{Where do people go when they die?} To the waves. Bodies are placed in boats and set adrift, weighted with stones carved with their name. If the boat sinks before the horizon, the dead is accepted. If it returns to shore, the dead is cursed. The beaches are littered with their bones.

\textbf{How do you apologize for a serious wrong?} You offer a storm-oath. Stand on a cliff during a gale and swear your apology to the wind. If the wind carries your words, you are forgiven. If it throws you into the sea, you are not.

\textbf{What is the most important festival?} The Launching of the Longships (spring thaw). The ships are dragged from their winter shelters and blessed with blood and mead. The ships are named after ancestors. The ancestors are listening. The ancestors approve.

\textbf{What is the secret that could destroy a powerful person?} The Volva of the Mist is a fraud. She cannot see the future. She reads the weather -- pressure, wind, the taste of salt on the air. The sailors know. The sailors do not care. A weather-reader is more useful than a seer.
\end{almanac}

\input{../generators/linns.tex}
\index{Linn}
% ======================================================================
% PART III: THE EASTERN DOLMIS
% ======================================================================
\part{The Eastern Dolmis}

\begin{quote}
  \small
  \textit{``East of the Dolmis, the sea grows saltier and the land grows stranger. The halflings count their blessings in thread and silence; the gnomes measure their debts in stone and courtesy. The forest remembers empires that crumbled to dust, and the pirate-city feeds on the bones of a sunken age. Each shore is a different kind of hunger -- but all of them remember the same ninth step.''}
  \\ --- The Edgewalker
\end{quote}

% -- Aelaerem --
\chapter{Aelaerem}
\emph{Hearth \& Hollow.}

\noindent
\textit{``The hollow counts your steps. The thread remembers your name. A proper harvest feast requires three things: the finest cider aged in the right barrel, a guest-loaf baked with intention, and the wisdom to know when to invite the quiet ones to sit by the fire. My geese know a stranger's step from a mile off, and they've never been wrong. If they start honking at noon, you best check your larder -- someone's coming who wasn't invited. And when the Hollow walks, the only safe place is a lit hearth with the ninth cup unpoured.''} \\
--- Apple-Matron Oria \textit{``Crimson Door''}

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Oatcakes and honey -- the oats are from the chalk downs, the honey from the bees that sleep in the barrow by the beech. The honey tastes of earth and old promises. The poor eat it with water; the rich eat it with cream and the knowledge that the Hollow might visit them instead.

\textbf{What does a child learn before they can walk?} To count in eights. Aelaerem children are taught the eight steps of the harvest dance, the eight knots of the guest-cord, the eight names of the ancestors who watch from the barrows. The ninth step is never taught. It is discovered -- or it discovers you.

\textbf{What is the first thing you notice about a stranger?} Whether they carry bread and salt. A stranger who offers bread and salt before crossing a threshold is a guest. A stranger who does not is a threat -- or a fool. The Hollow does not distinguish between the two.

\textbf{What is the most common crime?} Pouring the ninth cup. It is not a crime you can be arrested for. It is a crime the Hollow will punish. Those who pour the ninth cup find that the empty chair is no longer empty.

\textbf{What does no one talk about but everyone knows?} The Pale Shepherd is not a myth. He comes when lambs are born and when people go missing. He has no face. His hat is a shadow. He does not bargain. He only collects.

\textbf{Where do people go when they die?} Into the barrows. The dead are not buried -- they are placed in the hollow hills with a cup of cider and a loaf of bread. The barrows are the Hollow's larder. The dead are not dead. They are waiting.

\textbf{How do you apologize for a serious wrong?} You tie a red thread around your wrist and offer it to the wronged party. If the thread is accepted, you are forgiven. If it is cut, you are not. If it is tied around your throat, you have made a very serious mistake.

\textbf{What is the most important festival?} The Harvest Moot (autumn). The Apple-Matron pours the first cup for the eldest, the second for the guest, the eighth for the fire. The ninth cup is for the empty chair. The chair is never truly empty.

\textbf{What is the secret that could destroy a powerful person?} The Apple-Matron has never poured the ninth cup. The Hollow has been waiting for her to slip. She has not slipped. She is the only person in Aelaerem who has never made a mistake. The Hollow is patient. The Hollow is very, very patient.
\end{almanac}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
In Amedell, the Aelaerem keep their bonds with bread, bells, and careful counting. Lanes run under ash and hawthorn; doors reddened with thread promise a seat and a story to any who step right. Their magic is housekeeping writ large --- thresholds groomed, names jarred like jam, the Hollow fed with silence and the ninth cup left unpoured. When the Neighbors stir beneath the barrows, the folk answer with kettle songs and lantern law, and the fields behave because someone asked them to. Hospitality here is not kindness---it is a covenant. Every meal is a promise, every gift a thread, every candle an argument against the dark.

Amedell lies south of the Aelinnel homeland of Aevrossa, separated by the rolling downs and the slow waters of the River Adder. The two small-folk peoples share trade in red thread, copper, and seasonal hospitality, though the Aelaerem keep a wary distance from the fae-touched gnomish courts. It lies on the southeastern coast of the Dolmis, although the Aelaerem are rarely out to sea beyond coastal fishing.

\vspace{2mm}
\textbf{What you notice first:} The smell of baking bread and old apples. The way every threshold has a red thread tied at eye level. The silence of children when a stranger passes the village stone.

\textbf{The one rule that kills newcomers:} Never refuse an offered cup. The Hollow counts refusals as invitations. Pour the ninth cup for the empty chair, and you may feed it---or feed it.
\end{tcolorbox}

\subsection*{Regional Mood: Warm Hearths, Cold Thresholds}

Here, the land is not wilderness---it is a living weave of favors and omissions. The Aelaerem do not build walls; they build hedges, stiles, and the careful geometry of guest-right. A red door means bread, salt, and one safe night if you come honest. A cup-mark stone on the verge accepts butter as a prayer. The watch-geese know your step before you do, and the scarecrow turns to face the road when the Hollow is hungry. The Neighbors---the things that live under the barrows, in the uncounted ninth step, in the space between the candle and its shadow---are not enemies. They are neighbors. And neighbors, if treated politely, will stay on their side of the hedge.

\subsection*{The Hollow's Attention (Campaign Timer)}

The Hollow is not a monster. It is a forgetfulness that lives in thresholds, miscounts, and broken guest-right. Track its attention with a visible timer:

\textbf{Hollow Attention [8]} -- Advances when:
\begin{itemize}
\item A character refuses an offered cup (+1)
\item A promise is broken on a threshold (+1)
\item The ninth cup is poured (+2)
\item A guest-right is violated (GM spends 1 SB to tick +1)
\item A child points at an empty chair (the Hollow is already listening; +0, but the GM gains 1 SB)
\end{itemize}
At 4 segments: Lanterns flicker without wind. Dogs whine at nothing. The Hollow is watching.  
At 8 segments: The Hollow manifests. Roll on the \textbf{Hollow Walks} table. The party must offer a story, a memory, or a name to appease it.

\subsection*{Hollow Walks (d6)}
When the Hollow Attention timer reaches 8, roll 1d6 for the manifestation:
\begin{enumerate}
\item A pale figure stands at the edge of the circle. It asks for a name. Give a name (living or dead) and it leaves. Refuse, and it takes your shadow (--1 die on stealth for a week).
\item The ninth cup pours itself. Whoever drinks from it forgets their greatest joy (lose 1 XP, refundable by a quest).
\item A red thread appears tied around every PC's wrist. It tightens when you lie. Cutting it costs 1 Corruption.
\item The geese fall silent. The Hollow is inside the circle. It demands a story that ends with a death. If the story is true, it leaves. If false, it takes a memory.
\item The Moot Oak bleeds sap. The Hollow speaks through it: ``One of you carries a promise unpaid. Settle it now.'' The one named must offer 1 Boon or 1 Supply. Refusal ticks the timer again immediately.
\item The empty chair is occupied by a figure in gray. It does not speak. It eats. If you serve it food, it leaves. If you ignore it, it eats your luck (all rolls are Desperate for the next scene).
\end{enumerate}

\subsection*{Guest-Right Violations (d6)}
When a PC breaks guest-right (refuses a cup, draws steel under a red door, insults the host), roll 1d6:
\begin{center}
\begin{tabular}{|c|l|}
\hline
\textbf{d6} & \textbf{Consequence} \\
\hline
1 & The host demands a price of 1 Supply (the cost of a broken loaf). \\
2 & The village stones turn inward. You are not welcome anywhere for a week (--1 Position on all social rolls). \\
3 & A red thread ties itself around your wrist. It tightens when you speak to an Aelaerem. Cut it with a silver knife (cost 1 Boon) or suffer --1 die on all interactions. \\
4 & The watch-geese chase you from the village. Gain a \textbf{Marked by Geese} condition (animals are hostile; +1 DV to all Beastcraft rolls). \\
5 & The Hollow notices you. Tick \textbf{Hollow Attention +2}. The empty chair will be waiting at your next camp. \\
6 & The Apple-Matron herself banishes you. Start a \textbf{Gate of Thorns [4]} timer --- when full, you cannot enter any Aelaerem village until you pay a blood-price. \\
\hline
\end{tabular}
\end{center}

\subsection*{Prominent NPCs of Aelaerem}
\label{sec:aelaerem-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Oria \textit{``Crimson Door''} & Apple-Matron of the Harvest Moot & Keep the Hollow at bay and the feasts true & She pours the first cup for the eldest, the ninth for the empty chair. Never takes the ninth herself. \\
\hline
Bram & Miller of the Watch-Geese & Protect the village from unseen strangers & His geese have names; one is ``Stranger,'' and it has never been wrong. \\
\hline
Celia & Hedge-Witch Midwife & Tie red thread and untie feuds before they fester & Her thread is from her own hair. She never cuts it---only knots it. \\
\hline
Drogo & Thresher-King (seasonal) & Lead the harvest rites and keep the sheaf's secret & Chosen by the last sheaf, which is always a little damp. \\
\hline
Alda & Lantern-Warden of the Downs & Trim the lamps and know which shadows are wrong & Lights them in a pattern that is a prayer. \\
\hline
Finnian & Mummers' Captain & Guard the masks and the laws of festival & The masks are not carved. They are grown. \\
\hline
\end{longtable}

\input{../generators/aelaerem.tex}
\index{Aelaerem}

\newpage

% -- Aelinnel --
\chapter{Aelinnel}
\emph{Stone, Bough, and Bright Things.}

\noindent
\textit{``Count the steps, weigh the words, and never trust a single telling. The law here flows like water over stone -- it finds its own path, and woe to those who try to dam it with paper and pretense. Our kilns tell us more than your books ever could. When the smoke curls left, the tide's with you. When it splits in three, best stay by your own hearth till it burns straight again. And never trust a court that speaks in only one voice -- the silence is always the second witness.''} \\
--- Lady of Thorns

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Porridge with honey and a pinch of salt -- the salt is from the tide-rift steps, the honey from the Moonwell basin. The salt tastes of memory; the honey tastes of forgetting. The balance is everything.

\textbf{What does a child learn before they can walk?} To count in prime numbers. Aelinnel children are taught the sequence: 2, 3, 5, 7, 11, 13, 17, 19. The ninth number is not taught. It is discovered -- and it is always a door.

\textbf{What is the first thing you notice about a stranger?} Whether they carry iron. Iron is not forbidden in Aelinnel; it is simply \textit{impolite}. A stranger who carries iron is making a statement. The statement is heard. The hearing is remembered.

\textbf{What is the most common crime?} Breaking a geas. A geas is not a law; it is a word that has grown teeth. Those who break a geas find that the word follows them, and the word is always hungry.

\textbf{What does no one talk about but everyone knows?} The Green Gate is not a gate. It is a mouth. It opens when it is fed. It is always hungry. The Aelinnel feed it with exact change.

\textbf{Where do people go when they die?} Into the barrow galleries. The dead are not buried; they are placed in stone chimes that sing their names in the wind. The ninth note is always a sigh. The sigh is the Hollow's answer.

\textbf{How do you apologize for a serious wrong?} You carve an oath-bead. The bead warms when the apology is accepted. If it cools, the apology is not. If it glows, the Hollow has heard you, and the Hollow is not pleased.

\textbf{What is the most important festival?} The Counting of the Stones (spring equinox). Every Aelinnel counts the stones of their threshold. Those who count correctly are blessed. Those who count incorrectly are visited. The ninth stone is always warm.

\textbf{What is the secret that could destroy a powerful person?} The Stone Prince's seal is a lie. It was carved by a half-fae, half-mortal artisan who did not understand the law they were writing. The seal is still warm. It is still writing.
\end{almanac}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Stone, Bough, and Bright Things},breakable]
The Aelinnel do not build cities. They \emph{grow} them -- from granite and hawthorn, from tide-polished bone and the shadow of a laugh. Their halls thread through sea-cliffs and root-galleries, their roads are counted by antler-posts that move at night, and their doors open only when you have forgotten why you came. Here, promises have \emph{weight} -- not the weight of ink on parchment, but the weight of a stone dropped into a still pool. The ripples go on forever.

To the north, the Valewood breathes; to the south, the downs of Amedell where the Aelaerem tie red threads to keep the Hollow at bay. The Green Gate sometimes opens onto halfling barrows -- a fact neither people speaks of aloud, because speaking aloud gives the gate permission to move.

No single crown rules the gnomish lands. Dozens of courts, seasonal pacts, and garden societies weave a shifting lattice of loyalty. Some Aelinnel keep the old fae treaties, speaking in two voices and never carrying naked iron; others barter in the Green Market with memories instead of coin, treating courtesy as a kind of lockpicking. Still others vanish into the deep woods, listening to stag-roads where the trees whisper gossip about the living. The gnomish heart is divided -- and that division is the only reason the fae haven't eaten them whole.

Craft and courtesy govern every traffic between mortal keeps and the courts under the hill. Gnomish stone-singers hum bridges into alignment, their melodies keeping the Green Gate at bay. Forester-wardens hammer copper nails into thresholds because iron, they say, \emph{startles the truth}. In markets under living roofs, a sentence can be folded like a napkin, and a memory can be wrapped for travel -- but never opened before the second sunrise. Paths sometimes shorten for those who keep good count, and lengthen for those who sneer at patterns. The fae are never rude -- but they are never, ever \emph{safe}.
\end{tcolorbox}

\subsection*{Prominent NPCs of Aelinnel}
\label{sec:aelinnel-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Thorn & Lady of Thorns, Sovereign of the Hawthorn Court & Remind mortals that courtesy is a leash, not a gift & She never repeats herself. Her silences are verdicts -- and her verdicts are always polite. \\
\hline
Caelum Viridant & Green Knight, Warden of the Stag-Roads & Test the worthy and trap the careless & His armour is woven from living boughs that bleed when he lies. He has never bled. \\
\hline
Eilidh Stone-Prince & Mortal ruler of the Hall of Aelinnel & Keep the old pacts indexed and the fae at arm's length & She speaks only in conditional tense. No one has heard a declarative sentence from her lips in seven years -- not even her cat. \\
\hline
Riven & Huntsman of the Moonlit Ride & Chase down obligations and those who flee the Green Gate & His horn is carved from a fossilised scream. He only blows it when a promise is nine days old -- and nine days old is when the Hollow starts counting. \\
\hline
Tallow & Charcoal-burner (commoner exemplar) & Read the smoke and warn the village before the fae come & He has never lost a sheep to the Hollow. He has lost three helpers who did not listen. Their shadows still work his kilns. \\
\hline
Sable & Green-Market Broker & Trade truths for trinkets, and never lose the margin & Her receipts are written on leaf-vein paper. They curl when the bargain is unbalanced -- and also when it rains, which is often, which is why she keeps a stone in her pocket. \\
\hline
\end{longtable}

\input{../generators/aelinnel.tex}
\index{Aelinnel}

\newpage

% -- Valewood --
\chapter{Valewood}
\emph{Empire Under Leaves.}

\noindent
\textit{``The forest remembers. Speak truth, step lightly, and never eat the fruit of a tree with a face. The forest here remembers empires that crumbled to dust, and its laws are older than your kingdoms. To walk these paths is to enter into covenant with powers that measure time in seasons, not years. The woods here got a mind of their own -- they'll lead you where you need to go if you ask right, but they'll just as soon lose you if you're carrying ill intent. Best know which is which before you wander off the path.''} \\
--- Hazel Queen Amalthea

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Berry porridge and mushroom tea -- the berries are from the star-road groves, the mushrooms from the breathing city's lower levels. The porridge tastes of memory; the tea tastes of what you have forgotten.

\textbf{What does a child learn before they can walk?} To name the trees. Valewood children learn the eight sacred groves: Hazel, Alder, Oak, Ash, Thorn, Yew, Willow, and the ninth -- the grove that has no name. The ninth grove is where the empire's heart still beats.

\textbf{What is the first thing you notice about a stranger?} Whether they are followed by a shadow. The forest gives shadows to those who walk lightly. It takes them from those who do not.

\textbf{What is the most common crime?} Name-theft. The forest remembers names, and it can be persuaded to forget them. Those whose names are stolen are not dead -- they are simply no longer themselves.

\textbf{What does no one talk about but everyone knows?} The Genesis Vats still hum beneath the Amber Ziggurat. The Lethai-thora who sealed them did not die. They became something else. Something that still remembers the equations.

\textbf{Where do people go when they die?} Into the roots. The dead are not buried; they are woven into the forest's memory. The trees speak their names in the wind. The names are always a warning.

\textbf{How do you apologize for a serious wrong?} You offer a name-bead. If the bead warms, you are forgiven. If it cools, you are not. If it shatters, the forest has judged you, and the judgment is final.

\textbf{What is the most important festival?} The Phasing of the Star-Roads (autumn equinox). The star-roads align with the empire's fallen towers. Those who walk them at the right hour may see the empire as it was. Those who walk them at the wrong hour may never return.

\textbf{What is the secret that could destroy a powerful person?} The Hazel Queen is not a queen. She is a memory of a queen, preserved in amber and rooted in the forest's heart. The memory is fading. The forest is forgetting her name.
\end{almanac}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
The Valewood is older than kingdoms, older than the maps that skirt its borders. Beneath its canopy lie the bones of a Lethai empire whose name is forgotten, its towers swallowed by vine and root, its star-roads cracked and phasing between worlds. The air hums with half-remembered songs of magic, and the ruins bleed their mysteries into root and stream. To walk the Valewood is to risk being claimed by it: as guest, prey, heir---or trespasser.

But the empire did not fall to war or plague alone. In their final centuries, the Lethai-thora --- the high elf archivists and sorcerers --- grew desperate to match the Astroegro Empire's arcane might. They turned to forbidden arts: flesh-shaping, soul-stitching, and the binding of patterns that should never have been woven. Deep beneath the Amber Ziggurat, in chambers sealed with runes that weep, they built the \textbf{Genesis Vats}. Here, elf and beast were combined, spirit and stone were fused, and the first beast-kin were born --- not as servants, but as weapons. The vats still hum. The failures, the half-finished, the screaming ones, were loosed into the forest when the laboratories collapsed. They bred true. They became the monsters that now stalk the star-roads: the redcaps, the thorn-walkers, the things with too many eyes.

The empire fell, yes. But something else came through the cracks. A scent of honey and rot. A laugh from a clearing that should not exist. The Lethai-thora sealed the deepest laboratories and burned their records. The few survivors spoke of a ``soft place'' in the world's fabric, a realm of endless twilight where the fae had always waited. They carved warnings into stone: \textit{Do not open the ninth gate. Do not sing the song that calls them.} The gates were sealed. The songs were forgotten. But the forest remembers. And in certain clearings, when the moon is wrong, the air smells of jasmine and grave-soil.

\vspace{2mm}
\textbf{What you notice first:} The smell of wet moss and old stone. The way the light filters through leaves like stained glass. The silence that falls when you speak a name too loudly.

\textbf{The one rule that kills newcomers:} Never eat the fruit of a tree that has a face. The forest does not forget those who take, and the first bite is always a promise.
\end{tcolorbox}

\subsection*{Regional Mood: Leaves, Oaths, and Echoes}

Here, the past is not past --- it phases in and out like moonlight on water. The Lethai-ar who remain are not conquerors but custodians, tending to laws written before the first human kingdom rose. The fae courts of Hazel and Alder play their endless games of courtesy and obligation, and the beast-kin speak in parliaments where every vote is a scent. Imperial shades march through walls that no longer exist, and the star-roads hum with geometries that twist the unwary into yesterday.

But the true horror lies beneath. The Lethai-thora's experiments left scars that never healed. The beast-kin are not a single people --- they are the descendants of those who survived the vats, their lineages mixed with wild creatures, their memories fragmented. Some still hear the whispers of the old laboratories, calling them to seek the sealed doors. Others have built a society of scent and scar, their laws a deliberate forgetting of what they once were. The monsters --- the true monsters, the ones with no parliament --- are the failed experiments: the redcaps, the thorn-walkers, the things that hunger for warmth. They are the empire's unfinished business.

And somewhere, in a place where the trees grow in impossible spirals, the soft place still breathes. The fae courts are its children. They do not speak of what lies beyond the ninth gate. They only smile, and their smiles have too many teeth.

\subsection*{The Ninth Taboo Manifestation}
\begin{tcolorbox}[colback=black!3,colframe=red!60!black,title=The Ninth Taboo,breakable]
In the Valewood, the ninth leaf to fall from any tree points to a hidden threshold. The wise catch it or pretend not to see. Those who follow the leaf's direction find a place that was not there a moment ago, and something old waits inside. The threshold may lead to a sealed laboratory, a fae court, or a clearing where the air smells of jasmine and grave-soil. It is always warm.
\end{tcolorbox}

\subsection*{Prominent NPCs of the Valewood}
\label{sec:valewood-npcs}

\begin{longtable}{|p{0.2\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.15\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumored Quirk} \\
\hline
Queen Amalthea & Hazel Queen, sovereign of the Hazel Court & Preserve the old pacts and keep the empire from waking & She has never spoken a lie. She has never needed to. \\
\hline
King Thorn & Alder King, ruler of the Twilit Court & Gather old promises and expand the forest's reach & His crown is a noose of thorns. He wears it proudly. \\
\hline
Pathweaver Sila & Lethai-ar guide (commoner exemplar) & Knot safe routes and keep the star-roads from bleeding & Her cord has eight knots; the ninth is the one she never ties. \\
\hline
Bramble & Fox-headed courier of the beast-kin & Deliver messages that cannot be written & It asks something in return. What it asks is always a secret. \\
\hline
The Huntsman & Antlered hunter of the Moonlit Ride & Claim those who break the forest's laws & His face is a stag's skull; his voice is the wind in the leaves. \\
\hline
Shardwright Jacob & Artisan who repairs the star-roads & Wake the ancient ways before they fade forever & Her chisel is a tooth from the first dragon. She stole it. \\
\hline
The Genesis Keeper (Ghost) & A Lethai-thora archivist who died sealing the vats & Prevent anyone from opening the sealed laboratories & She appears as a shimmer of heat, speaking in equations that no longer add up. \\
\hline
\end{longtable}

\input{../generators/valewood.tex}
\index{Valewood}

\newpage

% -- Zakov --
\chapter{Zakov}
\emph{Salt \& Serpent.}

\noindent
\textit{``The tide keeps a better tally than any scribe. The reef remembers every secret whispered above its depths. The tide here keeps better tallies than any scribe, and the reefs remember every secret whispered above their depths. Power flows like the currents -- those who understand them rule the waves. The Astroegro learned that. They are at the bottom now. The Kahfagians think they are different. They are not. The sea does not care about flags. It only cares about weight. The sewers and the sea both carry what folks want to forget, but they also remember everything. Best keep your business above water and your enemies guessing which dock you'll use next. The Salt Prince's crown floats, the Reaping grows in the dark, and the tide pulls for anyone with a sharp knife and a sharper tongue. Lately, the Gray Lady has been asking questions. She's not from here. She's from somewhere colder. And she's not the only one.''} \\
--- Salt Prince Lev

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Salt fish and black bread -- the fish is from the shallows, the bread from the Dregs Quarter's ovens. The fish tastes of brine; the bread tastes of the Reaping's patience. The poor eat it dry; the rich eat it with butter and the knowledge that the tide always turns.

\textbf{What does a child learn before they can walk?} To read the tide. Zakov children learn the eight phases of the tide -- Low, Flood, High, Ebb, and the four in between. The ninth phase is not taught. It is discovered when the water rises and does not fall.

\textbf{What is the first thing you notice about a stranger?} Whether they have a shadow. The Reaping takes shadows first. A man without a shadow is not cursed -- he is already harvested. His body just hasn't stopped moving yet.

\textbf{What is the most common crime?} False manifests. Every third ship carries cargo that doesn't appear on its papers. The Syndicate knows. The Syndicate takes a cut. The Syndicate calls it a ``tariff surcharge.''

\textbf{What does no one talk about but everyone knows?} The Salt Prince is not the ruler. The Reaping rules. The Reaping sets the prices. The Reaping decides who drowns. The Reaping is the city's memory. The memory is the Reaping's.

\textbf{Where do people go when they die?} Into the deep. Bodies are weighted with chain and dropped in the Serpent's Spine. The reef is the city's archive. The archive is never balanced.

\textbf{How do you apologize for a serious wrong?} You offer a weight -- a debt paid to the Syndicate. The weight is calculated by the wronged party's estimate of loss. The silver is melted and poured into the reef. The reef remembers the weight.

\textbf{What is the most important festival?} The Turning of the Tide (spring equinox). The city's boats are blessed with brine and blood. The boats are named after ancestors. The ancestors are listening. The ancestors are not pleased.

\textbf{What is the secret that could destroy a powerful person?} The Salt Prince's crown is a cage. The man who wears it is not the same man who took it. The crown is warm. It hums. It is the Reaping's collar.
\end{almanac}

\begin{tcolorbox}[colback=black!3,colframe=blue!60!black,title={Theme \& Atmosphere},breakable]
Zakov is the pirate-city of the Dolmis coast, built on the drowned bones of an \textbf{Astroegro} port city that sank beneath the waves when that ancient empire fell. The old stone ruins --- shattered columns, flooded plazas, sealed vaults --- lie only a few fathoms down, and at low tide you can see their roofs breaking the surface like the spines of dead leviathans. The city's streets are built on the rubble, its piers driven into the wreckage, and its deepest cellars open into sunken chambers where the water is warm and still. The Astroegro are gone, but their secrets remain: crystalline shards that hum with a terrible patience, sealed doors that weep brine, and the slow, creeping corruption that the locals call the \textbf{Reaping}.

Zakov's streets are piers, its courts are taverns, and its rulers are whoever the Syndicates fear most this season. The Salt Prince claims the throne of driftwood and extortion, but the real power ebbs and flows with the tide --- and the tide remembers everything. Here, every manifest is a lie, every harbor a doorstep to a dozen conspiracies, and every fog-light a lure or a warning. The city breathes with the sea: at low tide, secrets exposed; at flood, the drowned walk the Sunken Quarter and knock on doors that should be locked. Zakov is not a place you visit. It is a place you survive --- or rule.

Above the water, three powers circle. \textbf{Naval Intelligence} watches from the \textit{Ember}, a Kahfagian cutter permanently anchored in the smuggler's gate, waiting for the Syndicates to make a mistake. \textbf{Patron Alaric} funds the Oligarchs' artifact hunts from a counting-house in the Crimson Docks, a man whose glass eye records everything it sees. And \textbf{the Salt-Bitten}, a cult that worships the Reaping's hunger, grows in the sewers and the sunken cellars, fed by the Ordo Artificum's scholars and the Oligarchs' greed.

Beneath it all, the \textbf{Reaping}. Centuries ago, the first Salt Prince --- then a common adventurer --- dove into a submerged Astroegro vault off the coast. What he found there no one knows for certain. Some say it was a living crystal, a fragment of a god that fell into the sea. Others whisper it is the father of Khemesh, the first hunger. The oldest tales call it simply \textit{the Reaping}: a hunger that does not kill, but \textit{harvests}. The Reaping seeps into Zakov like groundwater, corrupting slowly, tainting every soul who stays too long. Dreams curdle into obsession. Generosity rots into expectation. The city's famous ``everyone owes everyone'' is not a culture; it is a symptom. The Salt Prince keeps the Reaping's source sealed beneath his palace, but the seal is cracking, and the crystal grows.

\vspace{2mm}
\textbf{What you notice first:} The smell of brine and tar. The constant slap of water against stone. The way every tavern has a door that leads to a different dock --- and every dock has a name that changes with the tide. The faint, harmonic hum that comes from the deep, like a bell that never stops ringing. And the Gray Lady's mast, visible from every pier in the harbor, her steel hull dark against the salt-crusted wood.

\textbf{The one rule that kills newcomers:} Never take a coin that tastes of brine. The Serpent's Mark is not a blessing. It is a weight that drowns. And never look into a still pool at low tide. The Reaping looks back. The Gray Lady's agents look too.
\end{tcolorbox}

\subsection*{Regional Mood: Salt, Serpent, and the Crystalline Deep}

Here, the sea is not a border --- it is a landlord. Zakov rose from the wreckage of a sunken Astroegro port, and every stone is a salvage claim. The Salt Prince rules from a throne of driftwood and extortion, but the real lords are the Syndicate factors who own the docks, the eel-cults that whisper from the sewers (descended from Astroegro cultists who worshipped the deep), and the fae courts that launder their influence through tide-bargains. Ships arrive with goods no customs post will name, and depart with cargoes no manifest will show. The city's law is the Tide Timer : low water for killing, flood water for trading, slack for negotiation. Break the rhythm, and the city will break you.

Above the tide, a quieter war is fought. \textbf{The Kahfagian Admiralty} has known about Zakov's secrets for decades, and the Ordo Artificum, its secretive arcane division, has confirmed the Reaping's nature as a failed Astroegro containment system. Naval Intelligence, operating through deniable agents and cutouts, moves to contain the threat. \textbf{The Oligarchs}, funded by Patron Alaric, hunt the Astroegro artifacts that could control the Reaping --- or weaponize it. They fund the Syndicates, the Salt-Bitten, and even Naval Intelligence itself, playing every side to ensure they profit from the chaos. The agents of the Gray Lady and the Oligarchs' factors watch each other across the tavern tables of Zakov, trading secrets and bullets.

The Reaping is not a myth. It is a presence. It manifests as a crystalline growth that spreads through the submerged ruins, leaching color from stone and warmth from water. Those who stay in Zakov too long begin to dream of drowning; their shadows grow sharp and angular; their bargains feel heavier. The Reaping does not demand worship. It demands \textit{attention}. And it has been waiting for centuries. The Ordo Artificum's scholars believe they can control it. The Salt-Bitten believe they can feed it. They are both wrong.

\subsection*{The Reaping --- A Campaign Timer}
\label{sec:zakov-reaping}

The Reaping is not a monster to be fought. It is a pressure, a corruption, a slow unmaking of the soul. Track the party's entanglement with it using a \textbf{Reaping's Corrosion timer [8]}. Tick the timer when:
\begin{itemize}
  \item A character accepts a bargain from the Salt Prince or a Syndicate factor.
  \item A character uses a \textbf{Serpent's Mark} (see Diamonds).
  \item A character fails a Resolve test while in Zakov (drowning dreams, crystalline whispers).
  \item The party lingers in the Sunken Quarter or the Serpent's Spine for more than a scene.
  \item The party handles an Astroegro artifact without protective measures.
  \item The party accepts aid from Naval Intelligence, the Oligarchs, or the Salt-Bitten.
\end{itemize}
\noindent At \textbf{4 segments}: the character gains the \textit{Crystalline Whisper} condition (you hear the Reaping's voice in quiet moments; --1 die to resist fear). Their shadow begins to shimmer faintly. At \textbf{8 segments}: the character becomes \textbf{Harvested}. Their shadow detaches and walks into the sea, and they owe the Reaping a name (a true name of someone they love, or their own). They may still act, but they are marked. The GM may spend 2 SB to have a \textbf{Collector} --- a crystalline spirit in the shape of a drowned sailor --- appear and demand payment. To clear the timer, the party must:
\begin{itemize}
  \item Offer the Reaping something it values more than a soul: a memory of a crime no one witnessed, a true name from the Astroegro era (found in the sunken vaults), or a shard of its own prison (a heist into the Salt Prince's vault).
  \item Perform a deep-dive ritual at the Serpent's Spine, cutting a crystalline growth from the reef and replacing it with a lock of hair from a willing volunteer.
  \item Deliver an Astroegro artifact to the Ordo Artificum for ``containment'' --- though this rarely ends well for the artifact.
\end{itemize}

\subsection*{Geographic Integration}

Zakov sits at the cultural and maritime crossroads of the eastern Dolmis. To the north, the \textbf{Valewood}'s mist-shrouded coasts send driftwood and nightmares down the current; the Reaping's crystalline hum is said to be an echo of the forest's deepest hungers. To the south, the \textbf{Theona} isles regard Zakov with wary tolerance --- the Ninth Taboo keeps the Green Host at bay, but Zakov's smugglers know which coves avoid the Host's riders. Across the Dolmis to the east, the gnomish \textbf{Aelinnel} of Aevrossa trade fine copper and stranger goods through Zakov's black market, their counting-songs an eerie counterpoint to the Reaping's harmonic pulse. The halfling \textbf{Aelaerem} of Amedell keep a wary distance, their red-thread wards flickering when the crystal grows strong.

The city's lifeblood is the narrow sea-lanes between the Dolmis and the Amaranthine. \textbf{Thepyrgos} controls the straits with its chain-towers and bell-warning systems, while \textbf{Kahfagian} patrols enforce tariffs on the open water. Zakov's pirates and smugglers must bribe, bluff, or blast past both to reach the rich southern waters. The \textbf{Linns} of the northern fjords sometimes portage their longships south from the Aberderrin into the Dolmis, trading amber and walrus ivory for Zakov's more questionable cargoes --- and leaving behind storm-oaths that the local Syndicates are careful not to break. The \textbf{Kahfagian} presence is the newest and most dangerous: Naval Intelligence's \textit{Ember} is a permanent fixture, and the Oligarchs' factors are everywhere. The Syndicates have learned to tread carefully around the Gray Lady and her agents.

\subsection*{Prominent NPCs of Zakov}
\label{sec:zakov-npcs}

\begin{longtable}{|p{0.18\textwidth}|p{0.25\textwidth}|p{0.3\textwidth}|p{0.17\textwidth}|}
\hline
\textbf{Name} & \textbf{Role/Title} & \textbf{Motivation} & \textbf{Rumoured Quirk} \\
\hline
Salt Prince Lev & Nominal ruler of Zakov; keeper of the Reaping's seal & Keep the syndicates from tearing the city apart & His crown is a circlet of driftwood; he has never taken it off. Beneath it, his scalp is cracked like dry earth. Something crystalline moves in the cracks. He meets with Commodore Voss once a month. He always looks more tired afterward. \\
\hline
Pirate Queen Mila & Rules the Shallows with charm and cutlass & Carve out a kingdom free of the Salt Prince's levies & Her cutlass was her grandmother's; the grandmother was also a queen. The grandmother was the first to see the Reaping and refuse it. The refusal cost her her shadow. She does not know her grandmother was a Kahfagian agent. She does not know she is one too. \\
\hline
Commodore Voss & Commander, Kahfagian Naval Intelligence (Zakov station) & Contain the Reaping; recover Astroegro artifacts; keep the Syndicates in check & Her monocle is a polished Astroegro crystal shard. She sees the Reaping's corruption on anyone nearby. She has never taken it off. She sleeps with it on. She has not slept in three years. \\
\hline
Patron Alaric & Oligarch factor; head of the Zakov artifact trade & Acquire Astroegro artifacts; control the black market; feed the Ordo Artificum's research & His left eye is a functional Astroegro recording device. Everything he sees is archived. He has sold copies of his recordings to the Syndicates, to Naval Intelligence, and to the Salt-Bitten. He knows where the Reaping's source is. He is not telling. \\
\hline
Kraken's Tongue & Tide-oracle who speaks in riddles & Survive another season without being harvested & She lives in a bell tower that only exists at high tide; the bell does not ring. The bell is a fragment of the Reaping's prison. She tends it with her own blood. Her blood is crystalline. It glows in the dark. She has spoken to Commodore Voss once. She refused to tell her what she saw. \\
\hline
Scribe Marlo & Ordo Artificum scholar & Escape the Oligarchs' influence; expose the Ordo's true purpose & His hands are stained with crystal dust from handling artifacts. The dust has begun to grow. He has begun to dream in crystalline. He is afraid of what he will become. He knows the Ordo Artificum is not studying the artifacts. They are \textit{feeding} them. \\
\hline
The Salt-Bitten Queen & Leader of the Reaping cult & Awaken the Reaping; become its harvesters & She has no eyes --- only crystalline sockets. She can see the Reaping's future. She is leading her followers into the Sunken Quarter to build a crystalline cathedral. She has made a deal with Patron Alaric. He will not speak of it. \\
\hline
Dock-Master Olena & Syndicate factor who ``loses'' ships & Own the waterfront, one weight at a time & Her tally has a secret page that is a map of debtors' bones. The bones are crystalline. They move. She has sold information to Commodore Voss three times. She has sold information to Patron Alaric twice. She has not decided which side she is on. \\
\hline
The \emph{Ember}'s Bosun & Adjutant agent, Kahfagian Naval Intelligence & Gather intelligence; maintain the Admiralty's presence & Her left arm is prosthetic, made from salvaged Astroegro alloy. It never tarnishes. It hums at low tide. She has been in Zakov for twelve years. She does not remember what she was before. She does not want to. \\
\hline
\end{longtable}

\input{../generators/zakov.tex}
\index{Zakov}


% ======================================================================
% PART IV: THE SPACES BETWEEN
% ======================================================================
\part{The Spaces Between}

\begin{quote}
  \small
  \textit{``The Ways Between do not measure miles. They measure choices. A wrong turn might lead to yesterday. A right turn might lead to a version of yourself who made a different decision -- and is still alive to regret it.''}
  \\ --- The Edgewalker
\end{quote}

% -- Ubral --
\chapter{Ubral}
\emph{The Stone Between Spears.}

\noindent
\textit{``Ubral is a land of upland mists, stone cairns, and clan oaths held tighter than iron. Caught between the Aelerian mountains and the lowland courts of Viterra, its hills are scarred by old raids and crowned by watch-fires. Law is written not in charters but in cattle, blood-price, and songs. A guest's word can buy more than a sword, but one broken vow can spark a feud lasting generations.''} \\
--- Lady Isolde of the Scar

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Oatcakes and skyr -- the oats are grown in highland clearings, the skyr is cultured in stone crocks buried in the peat. The peat remembers the old kings. The old kings are buried beneath it.

\textbf{What does a child learn before they can walk?} To count cattle. Ubral children are taught the eight breeds of the hills. The ninth breed has no color -- it belongs to the cairn-folk. The dead do not pay taxes.

\textbf{What is the first thing you notice about a stranger?} Their guest-token. A token given by their last host. A warm token means recently welcomed. A cold token means traveling long. No token means a spy. Spies are given a rope.

\textbf{What is the most common crime?} Cattle raiding -- but raiding is a sport. A stolen cow is a stolen child. The return is a negotiation. The negotiation is a feud. The feud is a song.

\textbf{What does no one talk about but everyone knows?} The Lady of Tor is not the ruler. The cairns rule. The cairns vote at each full moon. Their votes are carried by the wind. The wind speaks to the hearth-aunts. The clan obeys.

\textbf{Where do people go when they die?} To the cairns. Their bones are added to the cairns. The cairns grow taller each year. The oldest cairn has no name. It remembers the first clan.

\textbf{How do you apologize for a serious wrong?} You offer a wergild -- payment in cattle. The cattle are driven to the cairn. If they graze peacefully, you are forgiven. If they scatter, you are not. If they refuse to enter the covenant's shadow, your claim is blood.

\textbf{What is the most important festival?} The Gathering of the Clans (autumn). The clans meet at the Great Cairn. The cattle are counted. The songs are sung. The feuds are settled. The dead are named. The hills remember.

\textbf{What is the secret that could destroy a powerful person?} The Great Cairn is a prison -- the tomb of a Vritann king who tried to conquer Ubral and failed. He was buried alive. His ghost votes with the cairns. He votes for war. The cairns have not yet agreed.
\end{almanac}

\input{../generators/ubral.tex}
\index{Ubral}

\newpage

% -- Silkstrand --
\chapter{Silkstrand}
\emph{City of Bridges \& Dyewater.}

\noindent
\textit{``Silkstrand is a city strung across canals and arched bridges, where every crossing is a negotiation and every current carries coin. The silk looms never sleep, the Exchange never closes, and the Archivolt's notaries seal fates with a drop of wax. The canals are not just water -- they are memory. Every bribe, every broken promise, every cargo never declared -- the canals remember.''} \\
--- The Bravo of Silkstrand

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Rice porridge with dried fish -- the rice is from Oshiiran barges, the fish from Zakov fleets. Both are smuggled. Both are delicious. Both are paid for in favors.

\textbf{What does a child learn before they can walk?} To count bridges. The eight bridges of the city: Merchant's, Weaver's, Dyer's, Cooper's, Pilot's, Beggar's, Saint's, and the ninth -- the bridge only there at low tide. The ninth bridge leads to the Sunken Quarter, where promises go to drown.

\textbf{What is the first thing you notice about a stranger?} Their ink. Every Silkstrand merchant has a seal-ink stain on their right thumb. Fresh ink means new to the trade. Faded ink means retired. No ink means not a merchant. Not a merchant is not a person.

\textbf{What is the most common crime?} Contract fraud -- but fraud is a negotiation. Contracts are written in disappearing ink. The notaries sell the antidote. The antidote is the fee.

\textbf{What does no one talk about but everyone knows?} The Matron is not the ruler. The canals rule. The canals set the prices. The canals decide who drowns. The canals are the city's memory. The memory is the Matron's.

\textbf{Where do people go when they die?} To the canals. Bodies are wrapped in silk and sunk in the deep channels. The channels are the city's archive. The archive is never balanced.

\textbf{How do you apologize for a serious wrong?} You offer a bridge -- a toll paid to the Bridge-Lords. The toll is calculated by the wronged party's estimate of loss. The silver is melted and poured into the canal. The canal remembers the weight.

\textbf{What is the most important festival?} The Dyeing of the Waters (spring). The dyers empty their vats into the canals. The water turns red, blue, green, gold -- the year's fashions. The Matron's favorite color is the one that sells best. It is also the color she dyed her own hair.

\textbf{What is the secret that could destroy a powerful person?} The canals were dug by a Sidhi engineer paid in promises. The promises were broken. The engineer cursed the city. Every ninth drop of water is a tear. She has been crying for nine centuries.
\end{almanac}

\input{../generators/silkstrand.tex}
\index{Silkstrand}

\newpage

% -- Midh-Akhaz --
\chapter{Midh-Akhaz}
\emph{The Violet Gate.}

\noindent
\textit{``Midh-Akhaz is not a city. It is a gap in the world where trade is the only law. No walls, no gates -- just the smell of horse sweat, cedar smoke, and the copper tang of coin paid in blood. The grass grows tall to conceal the fallen, not to honor them. Ecktorian emissaries call it ``the Brink'' -- the edge where coin buys survival. Caravan guards call it home -- a temporary respite from the weight of the road. Thieves call it paradise -- a place where the only law is the weight of a knife in the belt.''} \\
--- Aeler Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Curd cheese and flatbread -- the cheese is from Ykrul herds, the bread baked in dung-fire embers. The poor dip it in salt; the rich dip it in honey stolen from a border hive.

\textbf{What does a child learn before they can walk?} To count horses. The eight gaits of a healthy mount. The ninth gait -- the death-sprint -- is never taught. It is discovered when the rider's life depends on it.

\textbf{What is the first thing you notice about a stranger?} Their braids. Ykrul warriors braid with clan colors. Vilikari traders wear oiled braids with silver beads. A stranger whose braids are loose has just fought. A stranger with no braids is a slave or a ghost. Neither stays long.

\textbf{What is the most common crime?} Grass-right forgery. The Oasis Compact issues branded sticks for grazing rights. A clever carver can duplicate the brand. The penalty is exile into the tall grass at dusk. The grass has no court of appeals.

\textbf{What does no one talk about but everyone knows?} The Khatun is not the only ruler. The Silver Stall stable master knows where every stolen horse sleeps. The Dust Market auctioneer knows the price of every sin. The Iron Well keeper knows the name of everyone who will die before the next grass moon. They meet once a season. The grass records it.

\textbf{Where do people go when they die?} To the grass. Bodies are left beyond the caravan rings. The grass grows tall and green where a body falls. A patch of deep violet means the dead were not properly mourned. Locals avoid those patches.

\textbf{How do you apologize for a serious wrong?} You offer a blood-price on the Coin-Weigh Tribunal's scales. If the scale tips in your favor, you are forgiven. If it balances, you owe a story -- a true story that makes the wronged party weep. If it does not, you owe a finger.

\textbf{What is the most important festival?} The Grass Moon (autumn equinox). The violet seed-heads glow under moonlight. The tribes ride in a great circle, singing the names of the dead. At midnight, the Khatun pours water onto the earth. If it steams, the Hollow is walking. No one sleeps.

\textbf{What is the secret that could destroy a powerful person?} The Oasis Compact's original seal was stolen by a Vilikari factor who wanted to control water rights. The factor still lives in the Foreign Quarter under a false name. The seal is hidden in a hollow horse skull buried under the Grass Ring. The skull belonged to the Khatun's favorite stallion. The stallion did not die of old age.
\end{almanac}

\input{../generators/midh-akhaz.tex}
\index{Midh-Akhaz}

% ======================================================================
% PART V: THE WILDS, DUNGEONS, AND WAYS BETWEEN
% ======================================================================
\part{The Wilds, Dungeons, and Ways Between}

% -- Ways Between --
\chapter{The Ways Between}
\emph{Spiritways \& Veilways.}

\noindent
\textit{``The Ways Between do not measure miles. They measure choices. A wrong turn might lead to yesterday. A right turn might lead to a version of yourself who made a different decision -- and is still alive to regret it. The mist does not hide landmarks. It hides destinations.''} \\
--- The Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Nothing. There is no breakfast in the Ways Between. No morning, no hunger, only the walking. The walking is sustenance.

\textbf{What does a child learn before they can walk?} To forget. Children born in the Ways Between do not learn. They remember the path they took to be born -- eight turns, then the ninth turn is the birth. The birth is a choice. The child does not know they chose.

\textbf{What is the first thing you notice about a stranger?} Their direction. A stranger walking toward you may be coming from your future. A stranger walking away may be going to your past. A stranger walking beside you may be yourself from a timeline where you did not turn left. Do not ask.

\textbf{What is the most common crime?} Stealing -- but stealing is a paradox. A thief steals a memory from a version of you that no longer exists. The memory never belonged to you. You cannot miss it. You will miss it anyway.

\textbf{What does no one talk about but everyone knows?} The Ways Between are not between anything. They are the only thing that exists. The solid ground you remember -- cities, roads, hearths -- is the dream. The Ways are the waking. The waking is the walking. The walking never ends.

\textbf{Where do people go when they die?} Deeper. The Ways have no surface, no bottom. The dead walk deeper. The deeper they walk, the less they remember. When they remember nothing, they become the path. The path is the walking. The walking is the dead.

\textbf{How do you apologize for a serious wrong?} You cannot. Apologies require memory. Memory requires a self. The Ways erase both. The only apology is to stop walking. Stopping is not an option.

\textbf{What is the most important festival?} There are no festivals. There is only the moment between heartbeats when you realize you have been walking for nine years and you do not remember why. That moment is the festival. Celebrate it by taking the next step.

\textbf{What is the secret that could destroy a powerful person?} The Wayfinder is not a person. It is the road itself, dreaming of a traveler who will finally reach the end. The end does not exist. The road knows this. The road dreams anyway.
\end{almanac}

\input{../generators/ways_between.tex}
\index{Ways Between}

\newpage

% -- Wilds --
\chapter{The Wilds}
\emph{Roads, Ruins, and Weather.}

\noindent
\textit{``Between towns and treaties, the land breathes with patience. Old roads keep the step of armies in their dust; cairns hold the names that no page can contain. Law here is not written in books but in the marks we leave and the debts we owe to the earth -- not in coin, but in the weight of a story. To travel is not to conquer the land, but to ask its leave, and the land answers not with silver but with the memory of our passage. The hollows are not holes in the world; they are the lungs of the earth, breathing deep and slow. They do not judge; they do not give. They simply are, and in their being, they teach us to be.''} \\
--- The Bravo of Silkstrand

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Whatever they found -- berries, roots, a rabbit snared at dawn. The wilds provide. The wilds also take. The taking is breakfast for something else.

\textbf{What does a child learn before they can walk?} To read the sky. Wilds children are taught the eight winds by taste. The ninth wind has no direction -- it is the land's own breath. The land's breath is a warning.

\textbf{What is the first thing you notice about a stranger?} Their pack. Heavy pack means traveling long. Light pack means just started. No pack means a fool or a ghost. The wilds eat fools. Ghosts eat the wilds.

\textbf{What is the most common crime?} Forgetting -- but forgetting is a survival strategy. The wilds are full of things that want to be forgotten. A traveler who forgets them may pass safely. A traveler who remembers must pay a toll -- a story. The memory is the trap.

\textbf{What does no one talk about but everyone knows?} The wilds are not empty. They are full of watchers -- spirits, beasts, old things with no names. They do not interfere. They wait. They remember every footprint.

\textbf{Where do people go when they die?} Into the land. The dead are left where they fell. The land absorbs them. The land's memory grows longer. The dead become the land's memory. The memory is the wilds.

\textbf{How do you apologize for a serious wrong?} You leave an offering -- a coin, a button, a lock of hair. If the offering is gone by morning, you are forgiven. If it remains, you are not. If it has been moved, the land is testing you. The test is the walk. The walk is the apology.

\textbf{What is the most important festival?} The Turning of the Leaves (autumn). The wilds change color -- red is anger, gold is warning, brown is patience, green is forgetfulness. This year the leaves are red. The land is angry. Do not ask why.

\textbf{What is the secret that could destroy a powerful person?} The wilds have a heart. A stone buried beneath the oldest oak. The stone is warm and beating. If it stops, the wilds die. Nothing left to remember. Nothing left to wait.
\end{almanac}

\input{../generators/wilderness.tex}
\index{Wilds}

\newpage

% -- Dungeons --
\chapter{Dungeons}
\emph{Living Infrastructure.}

\noindent
\textit{``Beneath the surface, ancient places breathe with rusted lungs and dream in currents of forgotten speech. The walls do not forget a step; the air holds the echoes of words we wish we had not spoken. Dungeons are not broken things -- they are machines that have learned to live. Their lungs are the vents that sigh, their veins the pipes that throb, their jaws the doors that yield without touch. Some were forged by hand; others grew like mold in the dark. They are not ancient; they are patient, and they remember every flaw in the design as if it were a feature.''} \\
--- The Edgewalker

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Fungus bread and cave water -- the fungus grows on the walls, the water drips from the ceiling. Both taste of stone. Both are all there is.

\textbf{What does a child learn before they can walk?} To count footsteps. The eight sounds of the deep: Drip, Creak, Groan, Whisper, Scrape, Echo, Silence, and the ninth -- the dungeon's own heartbeat. The heartbeat is a door opening. The door is always behind you.

\textbf{What is the first thing you notice about a stranger?} Their shadow. Dungeon shadows are long, even in bright light. A shadow too short means they have been here too long. A shadow that moves independently means they have been touched by the deep. No shadow means they are not a stranger -- they are the dungeon's guest. The dungeon never lets guests leave.

\textbf{What is the most common crime?} Map-making -- but map-making is heresy. The dungeon changes its corridors when it feels watched. A map that is accurate at dawn is a lie by dusk. The punishment is to walk the lie until it becomes true.

\textbf{What does no one talk about but everyone knows?} The dungeon is not a place. It is a creature that dreamed itself into stone. The corridors are its veins. The chambers are its organs. The prisoners are its thoughts. The thoughts are not its own.

\textbf{Where do people go when they die?} Into the walls. Bones are ground into mortar and used to patch the cracks. The cracks always reappear. The dead are never finished being used.

\textbf{How do you apologize for a serious wrong?} You offer a memory -- a true one, witnessed by the dungeon itself. The memory is added to the walls. The walls whisper it back when you sleep. The apology never ends.

\textbf{What is the most important feature?} The Turning of the Keys (whenever the dungeon decides). Locks change, prisoners move, corridors rearrange. There is no date, no warning, no end. It is happening now. You did not notice. That is the point.

\textbf{What is the secret that could destroy a powerful person?} The dungeon's architect is still alive, sealed in the lowest chamber. The dungeon does not want to be alone. The architect is its only friend. The architect has not spoken in nine centuries. The dungeon does not need words.
\end{almanac}

\input{../generators/dungeons.tex}
\index{Dungeons}

% ======================================================================
% EPILOGUE
% ======================================================================
\chapter*{The Memory Closes}
\addcontentsline{toc}{chapter}{The Memory Closes}

\noindent
\textit{``The road is south. The ink is fresh. The lamp is borrowed.''} \\
\hfill --- Silkstrand Bravo

\vspace{0.5em}
\noindent
\textit{``The sea does not forgive, but it does forget -- if you give it a story worth remembering.''} \\
\hfill --- Lady Isolde

\vspace{0.5em}
\noindent
\textit{``The ledger is closed. The knife is still sharp. Your name is next.''} \\
\hfill --- Aeler Edgewalker

\vspace{2em}
\begin{quote}
  \small
  \textit{``The road is a memory. Every oasis is a story, every broken wheel a lesson. The Tulkani tie knots to remember who shared what. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. And the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.''}
  \\ --- Dusana of the Raven Road, at a caravanserai whose name she refused to give
\end{quote}

% ======================================================================
% CODA
% ======================================================================
\begin{chapter*}{Coda: A Random Encounter (Not Really) at the Three-Mugs Inn}
\addcontentsline{toc}{chapter}{Coda: A Random Encounter (Not Really) at the Three-Mugs Inn}

The night air was just shy of frost, and the Three-Mugs Inn smelled of spiced ale, damp wool, and the faint promise of gossip that had been sitting on the bar for a week. The fire crackled low, casting a wavering amber glow over the handful of patrons who pretended not to be listening to each other's stories.

A soft creak announced the door's opening. In stepped a woman whose scarred cheek caught the firelight like a fresh-drawn blade. She brushed snow from her cloak, set her sword-hand on the hilt of a well-worn blade, and surveyed the room with the practiced eye of someone who had counted the stones of every bridge she'd crossed.

``Evening,'' she said, voice low but carrying. ``I'm Isolde of the Scar. Does anyone know if the road north of the Belworth still has that\dots unsettling gap where the stones whisper?''

A ripple of laughter rose from the corner where a lanky figure leaned back, a grin playing at the corners of his mouth. His cloak was patched with bright dyed patches, and a dagger dangled from his belt like an after-thought. He tipped an imaginary hat.

``Name's\dots well, I don't really need one, but you can call me Bravo. I've been tracing the canals of Silkstrand all day---turns out the ninth bridge only appears when the tide's feeling generous. You looking for a shortcut, or just a story to trade for a mug?''

From the shadows near the hearth emerged a third figure, tall and gaunt, whose breath seemed to fog the air even as the fire licked his boots. He carried no weapon, only a small brass bell that hung from his neck, silent for now. His eyes were the color of stone-wet after a storm, and he spoke in a tone that made the crackling fire sound like a distant rumble.

``Aeler,'' he said, the name barely more than a whisper. ``I've been walking the old tunnels beneath the Aeler holds. The stone remembers every footfall, and it's been\dots restless lately. If you're looking for a place that won't collapse under the weight of your expectations, I know a few shafts that still hold.''

The trio fell into an uneasy silence, each measuring the other not with suspicion but with the quiet respect of strangers who have walked too many roads to be surprised by another traveler's tale.

Isolde glanced at the bravo's dagger, then at the bell at the Aeler's throat. ``You both carry\dots reminders.''

Bravo chuckled, tapping the hilt of his blade. ``A reminder that I still owe someone a drink.''

Aeler lifted the bell just enough for it to give a faint, reluctant chime. ``And a reminder that some debts are measured in breath, not coin.''

The innkeeper, a stout woman with a flour-dusted apron, slid three mugs across the bar without a word. The ale was dark, frothy, and smelled faintly of rosemary---exactly the sort of thing that made a warrior think of home, a rogue think of a clever bargain, and a stone-walker think of the weight of a promise kept.

They clinked mugs, the sound a brief, bright note against the low hum of the tavern. No grand quest was sworn, no ancient prophecy whispered. They simply sat, drank, and let the night settle into the familiar rhythm of strangers who, for a moment, shared the same fire, the same stale bread, and the same unspoken understanding that the road ahead would be just as muddy, just as bright, and just as full of stories as the one they'd just left behind.

When the moon rose high enough to silver the snow outside, each rose, tucked their respective token---sword, dagger, bell---back into their packs, and slipped back into the night. They did not exchange names again; they did not need to. The Three-Mugs Inn would remember them only as the three who came in, shared a drink, and left without making a fuss---exactly the sort of uneventful, perfectly ordinary encounter that makes a tavern worth its name.

\end{chapter*}

% ======================================================================
% INDEX
% ======================================================================
\printindex

\end{document}
