% !TeX program = lualatex
\documentclass[12pt,oneside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

% Suit symbols – used only in imported generator files, kept for compatibility
\newcommand{\spadesuit}{♠}
\newcommand{\heartsuit}{♥}
\newcommand{\clubsuit}{♣}
\newcommand{\diamondsuit}{♦}

% Almanac environment
\newenvironment{almanac}
  {\begin{quote}\small\itshape}
  {\end{quote}}

\begin{document}

% ----------------------------------------------------------------------
% FRONT MATTER
% ----------------------------------------------------------------------
\frontmatter

\title{\Huge Fate's Edge\\[0.3\baselineskip]
\Large The Way of Silk}
\author{Sarnai, Knot‑Tyer’s Hand \\ \small (apprentice to Caravan Judge Sarnai of the Nine Knots)}
\date{\today}
\maketitle

\tableofcontents

% ----------------------------------------------------------------------
% PREFACE – Framing device (Tulkani apprentice)
% ----------------------------------------------------------------------
\chapter*{Preface: The Knot‑Tyer’s Ledger}
\addcontentsline{toc}{chapter}{Preface}

\noindent
{\Large
\textit{My grandmother tied a knot for every caravan she led. When she died, her cord had nine hundred knots. I inherited the cord and the debt. She had promised to write this book. Now I must.}
}

The road is not a line on a map. It is a cord, and every traveller ties a knot. The Fhara tie knots in water – they call them wells. The Sidhi tie knots in light – they call them lighthouses. The Pereshi tie knots in fire – they call them coals. The Kuvani tie knots in grass – they call them remount stations. And we, the Tulkani, tie knots in memory – we call them stories.

I am Sarnai, Knot‑Tyer’s Hand. My grandmother was Caravan Judge of the Nine Knots. She travelled the Way of Silk from the Brass Coast to the eastern isles, from the Violet Steppe to the Crimson Valley. She promised to write a guide for those who would follow – a ledger of customs, strings, and tolls. She died before she could finish.

This book is the interest on her promise.

You will find no card‑drawing oracles here. The road does not speak in suits; it speaks in etiquette, in water shared, in lamps lit at dusk, in the ninth cup left unpoured. I have replaced the cartomancer’s compass with a caravan‑master’s toolkit: season wheels, mount tags, score types, and the cold arithmetic of standing and feud.

The peoples of the Way are not exotic. They are neighbours. They have their own laws, their own debts, and their own ways of remembering. The Fhara weigh water before silver. The Sidhi keep green flames burning for the dead. The Pereshi erase names they will not speak. The Kuvani read the sky in horse‑hoof echoes. And the Tulkani – we tie knots. When you cross a threshold, knock eight times. The ninth is for the Hollow.

I have tied a knot for every story in this book. The cord is heavy. The debt is not yet paid.

\vspace{1em}
\noindent\textit{— Sarnai, at a caravanserai whose name she is still learning to pronounce}

% ----------------------------------------------------------------------
% HOW TO USE THIS BOOK
% ----------------------------------------------------------------------
\chapter*{How to Use This Book}
\addcontentsline{toc}{chapter}{How to Use This Book}

This book is a companion to the \textit{Fate’s Edge Core Amaranthine Worldbook} and the \textit{System Reference Document (SRD)}. It provides cultural and procedural content for the eastern caravan routes – the Way of Silk – using a **travel‑and‑trade engine** instead of the card‑based Cartomancy chapter.

\subsection*{What You Need}
\begin{itemize}
  \item The \textit{Fate’s Edge SRD or Player/GM Guides} for core rules (attributes, skills, actions, Position/Effect, Clocks, Boons, Story Beats).
  \item A **Caravan/Tribe Sheet** (see Appendix) to track Standing, Feud, Exposure, Oath Ledger, and Notables.
  \item **Mount/Camel Tags** for desert, steppe, and highland travel.
  \item The **Season Wheel** to guide the campaign’s rhythm.
\end{itemize}

\subsection*{Regional Chapters}
Each chapter covers one of the five peoples of the Way:
\begin{itemize}
  \item An introductory paragraph (in Sarnai’s voice, biased but affectionate).
  \item A local quote.
  ítem An \emph{almanac} – a quick Q\&A about daily life.
  \item \textbf{Etiquette Hooks} – cultural actions that grant Position or Effect.
  \item \textbf{Strings} – tokens that act as leverage (water‑share tablets, freedman’s seals, fire‑coals, etc.).
  \item \textbf{Venue Tags} – modifiers for common locations (well‑courts, lighthouse courts, fire‑temples, remount posts, wagon circles).
  \item Then the full generator tables (Spades/Hearts/Clubs/Diamonds) imported from separate files.
\end{itemize}

\subsection*{Appendix: Caravan \& Tribe Rules}
This appendix distils the procedures from the \textit{Sands of Moon and Brass} expansion: Caravan/Tribe Sheets, Mount Tags, Season Wheel, Score Types (Desert Caravan, Oasis Siege, Council Moot, etc.), Weather/Heat Matrix, and a selection of generators (d66 targets, d12 hazards, council cases, verses & boons).

\vspace{1em}
\noindent\textit{“The road is a cord. Every knot is a promise. Untie it, and you pay the debt.”} – Sarnai’s grandmother, the first Caravan Judge

% ----------------------------------------------------------------------
% REGIONAL CHAPTERS
% ----------------------------------------------------------------------
\mainmatter

\chapter{Gateways \& Travel}
\emph{The road is a cord. Every knot is a promise. Untie it, and you pay the toll.}

\section*{Travel Modes of the Way of Silk}

\subsection*{Caravan Routes}
The Way of Silk is a network of seasonal caravan tracks connecting Fhara, Sidhi, Pereshi, Kuvani, and Tulkani territories. Travel is governed by the \textbf{Season Wheel}: spring for caravans, summer for heat survival, autumn for trade, winter for councils.

\subsection*{Desert Legs (Fhara)}
Oasis courts and well‑spirits. A \textbf{Water‑Share Tablet} or a \textbf{Drowned Coin} buys water. Never refuse a date offered by a well‑keeper – it is an oath of hospitality.

\subsection*{Coast Roads (Sidhi)}
Lighthouse courts and freedman’s seals. A \textbf{Lighthouse Charter} grants safe harbour. Always honour the \textbf{Bread \& Salt} ritual before negotiation.

\subsection*{Highland Roads (Pereshi)}
Fire temples and name‑erasure. A \textbf{Fire‑Shrine Token} or a \textbf{Rose from the Ruined Garden} pacifies the mountain spirits. Never step on a fallen book – written words carry the weight of law.

\subsection*{Steppe Legs (Kuvani)}
Remount stations and sky‑road songs. A \textbf{Remount Chain Mark} or a \textbf{Grazing Compact} buys fresh horses. Never ride a Kuvani horse without asking its name – it will throw you.

\subsection*{Wagon Circles (Tulkani)}
Itinerant courts and knot‑memory. A \textbf{Memory‑Bead} or a \textbf{Safe‑Haven Oath} guarantees sanctuary. Never sleep outside the circle at night – the Hollow walks.

\section*{Gateways to Other Expansions}

\begin{itemize}
  \item \textbf{Amaranthine Core (Kahfagia, Ecktoria, etc.)}: From Fhara, the \textbf{Salt Road} leads north to Oshiira. From Kuvani, the \textbf{Eastern Steppe Road} connects to Ykrul.
  \item \textbf{Southern Realms (Oshiira, Ashaan, Ngomebe)}: From Sidhi ports, sail to the Brass Coast (Dhahara) or to Oshiira. From Pereshi, the high passes lead to Ashaan.
  \item \textbf{Eastern Realms (Sihai, Nihori, Ayokha)}: From Kuvani, the \textbf{Silk Road} east leads to Sihai. From Tulkani, the \textbf{Processional Roads} connect to Nihori’s shrine circuits.
  \item \textbf{Dhahara (Brass Coast, Nine Kingdoms, Himadri)}: From Sidhi, the \textbf{Brass Coast} is a direct sea route. From Kuvani, the \textbf{Himadri passes} lead to the iron mines.
\end{itemize}
\newpage
% -------------------- FHARA --------------------
\chapter{Fhara}
\emph{The Oasis Courts.}

\noindent
Sarnai says: \textit{“The Fhara measure water before they measure silver. Their wells have ears. Their coffee has three cups – the fourth is for the Hollow. I tied a knot for every well I drank from. My cord is heavy.”}

\noindent
The Fhara dwell on a sun‑baked peninsula where scattered oases and dry riverbeds connect the coast to the interior. Their cities are built around wells and date groves; their caravans are the arteries of the spice trade. Water is law – the well‑keepers are judges, and the courthouse is the shade of a palm. Coffee is not a drink; it is a ritual. Three cups: welcome, negotiation, contract. Refuse the third, and you have broken faith.

\vspace{0.5em}
\noindent\textit{“Never refuse a date offered by a well‑keeper. The fruit is sacred to their well‑spirits. Refusing it ends the conversation – and your access to water.”} \\
— Well‑Judge Fatima al‑Wadi

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Dates and thin coffee – the dates from the well‑keepers’ groves, the coffee brewed with cardamom and camel milk. The milk is sour. The sourness is for endurance.

\textbf{What does a child learn before they can walk?} To count cups. The eight cups of negotiation. The ninth cup is for the well‑spirit. The well‑spirit is always thirsty.

\textbf{What is the first thing you notice about a stranger?} Their water‑skin. Full means not travelled far. Dry means travelled too far. No water‑skin means either a well‑keeper or a fool.

\textbf{What is the most common crime?} Water theft – a declaration of feud. The well‑spirit records every theft in the aquifer. The verdict is thirst.

\textbf{What does no one talk about but everyone knows?} The wells are not natural. The first Fhara carved them and bargained with the sand: water for memory. The Fhara remember everything. They cannot forget. That is the curse.

\textbf{Where do people go when they die?} To the oasis. Buried in date groves, marked with a single stone. If the name is remembered, the dead rest. If forgotten, they walk.

\textbf{How do you apologize for a serious wrong?} You offer the ninth cup, poured from your own water‑skin. Placed at the well’s edge. If the spirit drinks, you are forgiven. If the water evaporates, you are not.

\textbf{What is the most important festival?} The Filling of the Wells (spring). Every well cleaned and blessed. The water‑clerks recite the names of the year’s debtors. If their reflection wavers, they are forgiven. If it holds, they must pay.

\textbf{What is the secret that could destroy a powerful person?} The water‑clerks’ ledger is a forgery. The original was lost in a sandstorm three generations ago. The clerks have been inventing water rights ever since. The wells remember the original. They are waiting.
\end{almanac}

\subsection*{Etiquette Hooks}
\begin{itemize}
  \item \textbf{Water First}: Offer water before any business → +1 Position on Petition or Broker in Fhara courts.
  \item \textbf{Three Cups, One Contract}: Complete the three‑cup ritual (welcome, negotiation, contract) before a deal → +1 Effect on the final bargain.
  \item \textbf{Date Offering}: Accept a date from a well‑keeper → gain a temporary String (Well‑Spirit’s Favour) usable once in that oasis.
  \item \textbf{Faux Pas (SB)}: Refusing a date lets the GM start \textit{Thirst Clock} [4].
\end{itemize}

\subsection*{Strings (sample)}
\begin{itemize}
  \item \textbf{Water‑Share Tablet} – proof of contribution; reroll a failed Survival test when finding water.
  \item \textbf{Coffee Cup (Royal Seal)} – +1 Position in Fhara courts.
  \item \textbf{Drowned Coin of Qalat al‑Madina} – safe passage from any well‑spirit once.
\end{itemize}

\subsection*{Venue Tags}
\begin{itemize}
  \item \textbf{Well‑Court}: Neutral ground; Lamp‑Oaths bind as durable Strings; breaking ticks Exposure +1.
  \item \textbf{Coffee Bazaar}: Market starts at +1 Position; on a 1, \textit{Lamplighter Gossip} flips an Audience tag.
\end{itemize}

\newpage
\input{../generators/fhara.tex}
\newpage

% -------------------- SIDHI --------------------
\chapter{Sidhi}
\emph{The Diaspora of the Sealed Coast.}

\noindent
Sarnai says: \textit{“The Sidhi carry their homeland in a brass seal. They light green flames to show the drowned the way home. I asked one freedman if he would ever return to Galanina. He touched his collar – a scar, not a seal – and said, ‘I never left.’”}

\noindent
The Sidhi are a people of the coast and the diaspora. Their ancestral homeland is Galanina, a rugged shore of limestone cliffs and ancient ports. The Ashaani magocracy enslaved them for generations, scattering families across the Amaranthine. Those who survived the Long Trail – a forced migration spoken of only in whispers and lullabies – carried nothing but the memory of salt and the names of the dead. Now they run ropeways, spice auctions, and port guilds. Their symbol is the lighthouse lamp: a green flame that shows the faces of the drowned, so that the dead may witness the living.

\vspace{0.5em}
\noindent\textit{“Never ask about ancestry. The Ashaani slavery raids are a fresh wound. Asking which clan someone belongs to is asking how their grandparents were captured, sold, or died.”} \\
— Lighthouse Judge of Galanina

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Salted fish and flatbread – the fish dried on the docks, the bread baked in communal ovens fuelled by driftwood. The driftwood carries the names of lost ships.

\textbf{What does a child learn before they can walk?} To read the lighthouse codes. The eight lamp‑flashes: Safe, Danger, Storm, Pirate, Customs, Quarantine, Home, and the ninth – the flash that signals a freedman’s return. The ninth flash is never used. It is a promise.

\textbf{What is the first thing you notice about a stranger?} Their seal. Every Sidhi freedman carries a brass seal – the mark of manumission. Worn smooth means free for generations. Sharp means just freed. No seal means not a Sidhi. Not a Sidhi is not a person.

\textbf{What is the most common crime?} Seal‑forging – not a crime, a necessity. The Admiralty demands proof of freedom. The forgers are Sidhi. The Sidhi do not prosecute forgers. They hire them.

\textbf{What does no one talk about but everyone knows?} The brass seals are not brass. They are made from melted chains – the chains that bound the first freedmen. The metal remembers the weight of slavery. The seals are warm. The warmth is the memory of wrists.

\textbf{Where do people go when they die?} To the lighthouse. Ashes scattered from the tower’s top. The wind carries them to the sea. The sea carries them to Galanina. The ancestors are waiting.

\textbf{How do you apologize for a serious wrong?} You offer your seal – your own mark of freedom. It is melted down and reforged into a chain. The chain is worn by the wronged party. When the wrong is forgiven, the chain is melted again. One day, there will be nothing left.

\textbf{What is the most important festival?} The Lighting of the Green Flame (winter solstice). The lighthouse lamps are cleaned and refilled. The green flame shows the faces of the drowned. The ancestors watch. They do not judge.

\textbf{What is the secret that could destroy a powerful person?} The brass seals are not all equal. Some were forged by the Admiralty itself – trap‑seals that mark the wearer as a spy. The spies do not know they are spies. Their reports go to the Admiralty. The Admiralty knows everything.
\end{almanac}

\subsection*{Etiquette Hooks}
\begin{itemize}
  \item \textbf{Bread \& Salt}: Share a simple plate before negotiation → +1 Position in Sidhi homes or guild halls.
  \item \textbf{Name the Line}: Give your family or mentor’s lineage when introduced → Audience: Cool → Warm with civic officials.
  \item \textbf{Guest‑Right Token}: Present a stamped guest‑right from a guild → DV –1 on the first Petition in that venue.
  \item \textbf{Faux Pas (SB)}: Cut Line – ignoring lineage or guest‑right protocol imposes Position –1 for the scene.
\end{itemize}

\subsection*{Strings (sample)}
\begin{itemize}
  \item \textbf{Lighthouse Charter} – right to anchor in any Sidhi port without inspection.
  \item \textbf{Freedman’s Seal} – proof of emancipation; glows when a slaver ship is near.
  \item \textbf{Archive Key} – opens one door in the Sealed Archive of Galanina.
\end{itemize}

\subsection*{Venue Tags}
\begin{itemize}
  \item \textbf{Pilots’ Quay}: navigation disputes settled by local charts; Clue +1 on routes with a token.
  \item \textbf{Lighthouse Court}: tolls and wreck law; Board & Brace DV –1 when acting as sworn rescue.
\end{itemize}

\newpage
\input{../generators/sidhi.tex}
\newpage

% -------------------- PERESHI --------------------
\chapter{Pereshi}
\emph{The Roof of the Way.}

\noindent
Sarnai says: \textit{“The Pereshi keep coals that never cool. They speak of a prince whose name they erased. I asked a fire‑keeper why they remember him if they erased him. She said, ‘We do not remember him. We remember the debt.’”}

\noindent
The Pereshi highlands are a vast, windswept plateau, cut by deep river valleys and crowned with snow‑capped peaks. Ancient fire‑shrines stand beside caravanserais; terraced vineyards climb the hillsides. The people are scholars, shepherds, and traders – proud, suspicious, and bound by a code of hospitality that demands three days of feasting for any guest, and three generations of vigilance for any insult. The fire in the temple is not just a flame. It is a witness. It remembers every oath, every treaty, every lie told in its light.

\vspace{0.5em}
\noindent\textit{“Never step on a fallen book. Written words carry the weight of law and the weight of witness. To step on a book is to trample a contract, a genealogy, a verdict. The spirits of the words will haunt you.”} \\
— Flame‑Keeper Zarifa

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Flatbread with honey and clotted cream – the honey from highland bees, the cream from goats that graze the temple steps. The goats are sacred. The goats are also delicious.

\textbf{What does a child learn before they can walk?} To read a contract. The eight clauses of the Covenant. The ninth clause is never written – it is the silence where the unspoken terms live. The ninth clause is the one that kills you.

\textbf{What is the first thing you notice about a stranger?} Their book. Every Pereshi carries a personal ledger of debts owed and collected. A thick book means many debts. A thin book means many creditors. No book means a liar. Liars are not fed.

\textbf{What is the most common crime?} Contract forgery – but forgery is an art. The judges are forgers themselves. The law is a negotiation that never ends.

\textbf{What does no one talk about but everyone knows?} The Forgotten Prince did not vanish. He is buried beneath the largest fire temple, in a lead‑lined coffin sealed with nine locks. The ninth lock has no key – the key is a memory no one wants to have.

\textbf{Where do people go when they die?} To the fire. Cremated on the temple pyres, ashes mixed with clay and pressed into bricks. The bricks build new temples. The dead become walls. The walls witness.

\textbf{How do you apologize for a serious wrong?} You offer a coal – a burning ember from the temple fire. Placed in the wronged party’s hearth. If it stays lit, you are forgiven. If it goes out, you are cursed. The curse is the fire’s silence.

\textbf{What is the most important festival?} The Reading of the Ledgers (autumn). Every household brings its ledger to the temple. The debts are tallied. The dead are named. The fire listens. This year, the ninth coal has not stopped flickering.

\textbf{What is the secret that could destroy a powerful person?} The ninth coal is not coal. It is a piece of the Forgotten Prince’s heart, petrified. It is still warm. It is still beating.
\end{almanac}

\subsection*{Etiquette Hooks}
\begin{itemize}
  \item \textbf{Fruit \& Tea}: Offer fruit and a warm cup on arrival → +1 Position in household negotiations.
  \item \textbf{Shoes at Threshold}: Remove shoes indoors → negate a Position –1 penalty from prior brusqueness.
  \item \textbf{Couplet of Courtesy}: A brief couplet acknowledging hosts’ craft → Audience: Cool → Warm with artisans/scribes.
  \item \textbf{Faux Pas (SB)}: Dust on the Mat – barging in tracks Exposure +1 among neighbours.
\end{itemize}

\subsection*{Strings (sample)}
\begin{itemize}
  \item \textbf{Fire‑Shrine Token} – +1 Position in any Pereshi court or negotiation.
  \item \textbf{Rose from the Ruined Garden} – burn to ask a yes/no question of the fire.
  \item \textbf{Tablet of a Severed Name} – contains an erased name; speak it to restore a person’s memory.
\end{itemize}

\subsection*{Venue Tags}
\begin{itemize}
  \item \textbf{Roof‑Garden Pact}: settlements on rooftops; a hosted pact counts as a String across partnered houses.
  \item \textbf{Archive Court}: citations from the Fire‑Temple archive lower DV by 1 once/scene.
\end{itemize}

\newpage
\input{../generators/pereshi.tex}
\newpage

% -------------------- KUVANI --------------------
\chapter{Kuvani}
\emph{The Remount Steppe.}

\noindent
Sarnai says: \textit{“The Kuvani do not ride horses. They borrow them. A horse’s name is its lineage. Its lineage is its story. I told a remount master a story about a wolf who forgot her pack. He gave me a grey mare. The mare knew the story already.”}

\noindent
The Kuvani ride the eastern steppe beneath the Eternal Blue Sky. Kuva, the sky father, sees all: every oath sworn in the open, every horse born, every blade of grass bent by the wind. They have no temples – the sky is their shrine. Their shamans, called Wind‑Singers, read the future in the flight of hawks and the whisper of grass. Their dead are not buried; they are given to the sky on funeral platforms, that Kuva may carry their souls to the ancestral herd that rides the northern lights. A horse is not a mount; it is a relative. Its name is a lineage stretching back to the first mare that Kuva breathed life into.

\vspace{0.5em}
\noindent\textit{“Never ride a Kuvani horse without asking its name. The horse will answer by throwing you. If it throws you, you have insulted its lineage.”} \\
— Remount Master Alp Arslan

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Fermented mare’s milk and dried meat – the milk from the herd, the meat from the hunt. The hunt is a ceremony. The ceremony is a debt to the grass.

\textbf{What does a child learn before they can walk?} To ride. Placed on a horse before they can stand. The horse is patient. It remembers the child’s grandfather.

\textbf{What is the first thing you notice about a stranger?} Their saddle. Worn means many miles. New means many miles ahead. No saddle means a ghost. Ghosts do not need saddles.

\textbf{What is the most common crime?} Horse theft – a declaration of feud. A stolen horse is a stolen relative. The feud lasts until the horse is returned or the thief’s line is ended.

\textbf{What does no one talk about but everyone knows?} The ghost herd is real – spectral horses that ride the salt pan on moonless nights. They are the souls of riders who died without honour. They ride forever, searching for a battle that will never come.

\textbf{Where do people go when they die?} To the sky. Placed on funeral platforms, wrapped in felt. The birds take the flesh. The wind takes the bones. The sky takes the soul.

\textbf{How do you apologize for a serious wrong?} You offer a horse – the best in your herd. If the wronged party accepts, the debt is paid. If they kill the horse, the debt is blood.

\textbf{What is the most important festival?} The Ride of the Ancestors (autumn equinox). The clans ride to the salt pan. The ghost herd appears at midnight. The living ride with the dead.

\textbf{What is the secret that could destroy a powerful person?} The Khagan’s horse is a ghost. It was killed in battle twenty years ago. The Khagan does not know. He rides a skeleton wrapped in hide. The horse does not tire, does not eat, does not forgive.
\end{almanac}

\subsection*{Etiquette Hooks}
\begin{itemize}
  \item \textbf{Guest Tent Rope}: Leave a rope coiled outside to signal neutrality → +1 Position on first Parley in camp.
  \item \textbf{Bowl Before Blade}: Accept a simple bowl before speaking of arms → Audience: Cool → Warm with elders.
  \item \textbf{Honour the Horse}: Praise mounts; tend tack before sleep → DV –1 on next Chase related to horse teams.
  \item \textbf{Faux Pas (SB)}: Spurned Mount – neglecting a horse lets the GM impose Position –1 among riders.
\end{itemize}

\subsection*{Strings (sample)}
\begin{itemize}
  \item \textbf{Remount Chain Mark} – a horse’s tooth; exchange for a fresh mount at any Kuvani station.
  \item \textbf{Grazing Compact} – sheepskin document; safe passage on clan land.
  \item \textbf{Sky‑Road Song} – a melody on a bone flute; navigate any storm.
\end{itemize}

\subsection*{Venue Tags}
\begin{itemize}
  \item \textbf{Remount Post}: swap fresh mounts; once/leg treat Chase (horse) DV –1.
  \item \textbf{Sky‑Road Cairns}: waystones; spend a song token to convert Navigate Position +1 across open grass.
\end{itemize}

\newpage
\input{../generators/kuvani.tex}
\newpage

% -------------------- TULKANI --------------------
\chapter{Tulkani}
\emph{The Itinerant Courts.}

\noindent
Sarnai smiles: \textit{“My people have no home but the road. Our wagons are our hearths. Our knots are our memories. I tie a knot for every story I collect. When the cord is full, I will return to my grandmother’s circle and untie them, one by one, until the debt is paid.”}

\noindent
The Tulkani have no fixed homeland; they are wanderers who follow the seasons and the trade routes. Their wagons are their hearths, their circles are their courts, and their knots are their memories. They are welcome everywhere and trusted nowhere – a position they wear with wry humour. A Tulkani caravan can appear at a lord’s border, camp for seven nights, and leave before the eighth – because on the eighth night, the lord can demand rent. On the ninth night, something else comes. Their neutrality is sacred: a Tulkani camp is sanctuary for any traveller, regardless of allegiance, as long as they respect the circle and offer a story when asked.

\vspace{0.5em}
\noindent\textit{“Never sleep outside the circle at night. The wagon ring is sanctuary. Outside it, the Hollow walks. A Tulkani will not rescue you – they will only mourn you at dawn.”} \\
— Caravan Judge Sarnai (my grandmother)

\vspace{1em}
\begin{almanac}
\textbf{What do people eat for breakfast?} Tea and hardtack – the tea brewed from leaves gathered along the road, the hardtack baked at the start of the season. The hardtack is always stale. The tea is always fresh.

\textbf{What does a child learn before they can walk?} To tie a knot. The eight journey‑knots: Departure, Arrival, Promise, Debt, Payment, Witness, Apology, and the ninth – the knot that is never tied. The ninth knot is for the Hollow.

\textbf{What is the first thing you notice about a stranger?} Their cord. Every Tulkani carries a knotted cord – their life’s ledger. Many knots mean travelled far. Few knots mean just begun. No cord means a ghost.

\textbf{What is the most common crime?} Story‑theft – but story‑theft is a compliment. A stolen story is a story worth telling. The original teller is honoured. The thief is welcomed.

\textbf{What does no one talk about but everyone knows?} The Tulkani were exiled from their homeland nine centuries ago, for a crime that no one remembers. The crime is written in the ninth knot of the Elder’s cord. The cord is too long to read.

\textbf{Where do people go when they die?} To the road. Bodies left at a crossroads. The road takes them to the next valley, the next telling. The dead become stories. Stories are immortal.

\textbf{How do you apologize for a serious wrong?} You untie a knot – one knot from your cord, chosen by the wronged party. The knot is burned. The ash is scattered on the wind. If the wind blows toward the wronged party, you are forgiven.

\textbf{What is the most important festival?} The Tying of the New Cord (new year). Every Tulkani receives a new cord, unknotted. The old cord is burned. The ash is mixed with water and drunk. The drink tastes of memory.

\textbf{What is the secret that could destroy a powerful person?} The Tulkani are not wanderers. They are pilgrims – walking toward a destination that no longer exists. Their homeland was swallowed by the desert. The desert is still hungry. They will never arrive.
\end{almanac}

\subsection*{Etiquette Hooks}
\begin{itemize}
  \item \textbf{Story as Kinship}: Tell a true story (in character) to a Tulkani elder → gain a temporary String (Memory‑Bead).
  \item \textbf{Circle Courtesy}: Do not step over the cord that marks the wagon circle → first social roll in the circle gains +1 Position.
  \item \textbf{Ninth Night Offering}: Leave a knotted cord at the central post → the Hollow’s attention is delayed by one night.
  \item \textbf{Faux Pas (SB)}: Counting to nine aloud – the GM may start \textit{Hollow’s Patience} [4] immediately.
\end{itemize}

\subsection*{Strings (sample)}
\begin{itemize}
  \item \textbf{Memory‑Bead} – a single knot; present it to ask for a story from any Tulkani.
  \item \textbf{Safe‑Haven Oath} – a knotted cord; guarantees sanctuary in any Tulkani circle.
  \item \textbf{Dream‑Cord} – a silk cord with nine knots (the ninth loose); sleep on it to ask Ikasha one question in a dream.
\end{itemize}

\subsection*{Venue Tags}
\begin{itemize}
  \item \textbf{Wagon Circle}: neutral ground; no violence permitted; breaking the circle starts \textit{Exile Clock} [4].
  \item \textbf{Caravan Court}: disputes settled by storytelling; a good story (Performance test DV 3) grants +1 Effect on the verdict.
\end{itemize}

\newpage
\input{../generators/tulkani.tex}
\newpage

% ----------------------------------------------------------------------
% APPENDIX: CARAVAN & TRIBE RULES
% ----------------------------------------------------------------------
\appendix
\chapter{Caravan \& Tribe Rules}

This appendix distils the travel‑and‑trade engine from the \textit{Sands of Moon and Brass} expansion. Use these procedures when running campaigns focused on the Way of Silk.

\section*{Caravan / Tribe Sheet}
\begin{longtable}{|p{0.3\textwidth}|p{0.55\textwidth}|}
\hline
\textbf{Caravan/Tribe Name} & \\
\hline
\textbf{Seat (home oasis/camp)} & \\
\hline
\textbf{Council Affiliation} & (Fhara, Sidhi, Pereshi, Kuvani, Tulkani – or independent) \\
\hline
\textbf{Strings (2–3)} & water‑share tablet ∙ lamp‑oath seal ∙ trade charter ∙ feud settlement ∙ way‑song ∙ guide‑stone \\
\hline
\textbf{Tags (2–4)} & Water‑Wise ∙ Star‑Pathed ∙ Heat‑Trained ∙ Merchant‑Known ∙ Hospitable ∙ Law‑Strict ∙ Poet‑Loud ∙ Falcon‑Banner \\
\hline
\textbf{Tracks} & Standing [6] (reputation), Feud [4] (active quarrel), Exposure [6] (to foreign powers) \\
\hline
\textbf{Oath Ledger} & favours owed/held (name + what; acts like Favour/Leverage) \\
\hline
\textbf{Notables} & guide ∙ war‑leader ∙ speaker ∙ poet ∙ water‑keeper \\
\hline
\end{longtable}

\subsection*{Standing [6]}
Tick up for kept oaths, fair trade, valour. Tick down for oath‑breaking, cowardice, sacrilege. High Standing grants \textit{Audience: Respectful}; low invites \textit{Audience: Skeptical}.

\subsection*{Feud [4]}
Name the counterparty. Ticks on insult, theft, or harm. Clear by blood‑silver, ordeal, or deed. When filled, blood‑price owed or open war.

\subsection*{Exposure [6]}
Ticks when the party attracts attention from foreign powers (Admiralty, Veiled Sisters, etc.). When filled, a hostile faction acts directly.

\section*{Mount / Camel Sheet}
Pick a mount and 3–4 tags.

\paragraph{Dromedary (caravan)}: water‑efficient, steady pace, cargo capacity.
\paragraph{Charger (war)}: fast sprint, agile turns, combat‑trained.
\paragraph{Sand Runner (scout)}: light, narrow feet, collapsible saddle for dunes.

\subsection*{Mount Tags (choose 3–4)}
\begin{itemize}
  \item \textbf{Sand‑Sure}: Desert sand Drive/Handle DV –1; dune ridges, quicksand.
  \item \textbf{Water‑Efficient}: Half water consumption; ignore first Fatigue from heat.
  \item \textbf{Night‑Runner}: In darkness, Navigate DV –1; Ambush at Dawn starts one Position higher.
  \item \textbf{Heat‑Ready}: In sun/fire scenes, first Condition +1 is ignored.
  \item \textbf{Qanat‑Swimmer}: Underground water passages; Disengage gains Position +1 in tunnels.
  \item \textbf{Falcon‑Eye}: +1 die on Lookout/Watches; spot distant smoke/sand columns.
  \item \textbf{Brass‑Bearer}: Carries heavy loads; ignore encumbrance penalties for trade goods.
  \item \textbf{Poet’s Drum}: Once/score gain \textit{Audience: Fierce} aboard; can rally morale.
  \item \textbf{Guide‑Stones}: Hidden cairn‑marks; once/score Navigate Position +1 in trackless sand.
\end{itemize}

\section*{Roles (choose at table)}
Guide (navigation), War‑Leader (combat/assault), Speaker (parley), Poet (morale/saga), Water‑Keeper (supplies/repairs), Lookout (weather/sandstorm).

\section*{Season Wheel}
\begin{itemize}
  \item \textbf{Spring}: Caravan season – trade routes open.
  \item \textbf{Summer}: Heat – survival focus; oases contested.
  \item \textbf{Autumn}: Trade/Settle – markets busy; councils meet.
  \item \textbf{Winter}: Council – disputes settled; oaths renewed.
\end{itemize}
At each transition: advance Politics (Mandate/Crisis) for regional powers; roll Weather/Heat [4–6] for the current theatre; offer Oath Opportunities (escort, feud settlement, shrine warding).

\section*{Score Types}
Pick approach (Deceit ∙ Speed ∙ Shock ∙ Parley) → set Position/DV from venue/tags → roll.

\subsection*{Desert Caravan (Profit / Passage)}
Clocks: Alarm [4–6], Trade [6–8], Mount Damage [4], Blood‑Price [4].  
Entry: sand gate, hidden wadi, qanat shortcut, market ruse.  
On 1s: GM spends SB → sandstorm warning, hidden quicksand, rival caravan ambush, guide betrayal.  
Resolution: When Trade fills, choose coin or Strings. If Blood‑Price fills, mark Feud +1 or pay blood‑silver.

\subsection*{Oasis Siege (Water / Tribute)}
Clocks: Thirst [4], Defense [6], Sentries [4], Terms [6].  
Position tweaks: Water‑Efficient/Heat‑Ready help; sandstorms/hail hurt.

\subsection*{Council Moot (Law / Settlement)}
Venue: circle of stones, winter pavilion, qanat meeting hall.  
Moves: Oath‑Swear (commit under penalty), Wager Wyrd (ordeal by feat), Blood‑Silver (compensation roll), Witness the Verse (Poet stakes a truth).  
Outcomes: resolve Feud, write Oath to ledger (durable String), assign seasonal water rights or trade route.

\section*{Weather / Heat Matrix}
Advance on 1s or fiction:
\begin{itemize}
  \item \textbf{Khamsin/Sandstorm}: sight Position –1; Sand‑Sure cancels. On 1, Pursuit +1 (lost bearings).
  \item \textbf{Flash Flood}: ranged actions –1 die; on 1, Condition +1 (saddle/gear).
  \item \textbf{Ward Storm}: Navigate DV +1; on 1, choose Delay (Distance stalls) or Sand‑Blast (Condition +1).
\end{itemize}

\section*{Generators (Quick Reference)}
\subsection*{Desert Targets (d66)}
11 star‑temple ∙ 12 beacon tower ∙ 13 cliff monastery ∙ 14 salt works ∙ 15 oasis fleet ∙ 16 lord’s pavilion ∙ 21 dune storehouse ∙ … (full table in \textit{Sands of Moon and Brass})

\subsection*{Desert Hazards (d12)}
1 dune on a bend ∙ 2 sudden flash flood ∙ 3 quicksand pool ∙ 4 deadwood snag ∙ 5 scorpion nest ∙ 6 heat shimmer ∙ 7 hidden side‑cut ∙ 8 rock echo ∙ 9 toll chain half‑raised ∙ 10 reeds conceal archers ∙ 11 rain‑swollen ford ∙ 12 sand‑suck bank.

\subsection*{Council Cases (d12)}
1 insult in verse ∙ 2 stolen guide stone ∙ 3 broken oath on winter grain ∙ 4 blood‑price disputed ∙ 5 marriage claim ∙ 6 way‑song stolen ∙ 7 hostage pledge lapsed ∙ 8 salvage rights ∙ 9 border cairn moved ∙ 10 feud cooling terms ∙ 11 verse witness contest ∙ 12 mercenary pay withheld.

\subsection*{Verses \& Boons (d12)}
1 drum‑song that steadies hands ∙ 2 falcon omen at dawn ∙ 3 amber find ∙ 4 guide’s ghost shows a cut ∙ 5 sand‑storm spares you ∙ 6 omen of red sails ∙ 7 poet’s verse spreads ∙ 8 winter pavilion adopts you ∙ 9 qanat seal renewed ∙ 10 cliff bell silent ∙ 11 stars open path ∙ 12 oath‑ring warms (truth told).

% ----------------------------------------------------------------------
% EPILOGUE
% ----------------------------------------------------------------------
\chapter*{Epilogue: The Ninth Knot}
\addcontentsline{toc}{chapter}{Epilogue}

My grandmother’s cord had nine hundred knots. I have added nine – one for each story in this book. The cord is still heavy. The debt is not paid. But I have tied the ninth knot loosely, as the Tulkani do. The Hollow is patient. So am I.

On the road to the Crimson Valley, with a fresh notebook and a borrowed lamp. The ink is squid. The wind is north. The next tale is already waiting.

\vspace{1em}
\noindent\textit{— Sarnai, Knot‑Tyer’s Hand}

\vspace{2em}
\begin{quote}
  \small
  \textit{“The road is a ledger. Every oasis is a credit, every broken wheel a debt. The Tulkani tie knots to remember who owes whom. The Fhara weigh water before they weigh silver. The Pereshi recite verses to keep the bridges open. And the Sidhi keep lighthouse lamps burning for the dead who still have not reached the shore.”}
  \\ — Dusana of the Raven Road (from whom I learned the telling)
\end{quote}

% ----------------------------------------------------------------------
% COLOPHON
% ----------------------------------------------------------------------
\chapter*{Colophon}
\addcontentsline{toc}{chapter}{Colophon}

This book was written on paper that had been a cargo manifest, a love letter, and a bandage. The ink is lampblack cut with salt water and the ash of a burnt confession. The binding is grey cord tied with nine knots. The ninth knot is not cut.

Sarnai composed the preface in a Tulkani wagon circle, while a Oath‑Keeper counted her breaths. The almanacs were gathered from traders, spies, and the occasional honest farmer. The generator tables were tested in the field – meaning she ran them at gunpoint in a caravan court while a Fhara well‑judge weighed her words.

If you find yourself on the Way of Silk and a Tulkani offers you a cup of tea, accept it. If they offer you a knotted cord, think carefully before you untie it.

The road is a cord. Every knot is a promise. Untie it, and you pay the debt.

\end{document}
