% ============================================================
% Fate's Edge --- Clean Preamble (pdflatex safe)
% ============================================================

\documentclass[11pt,letterpaper,twoside]{book}
% ---------- Encoding, fonts, micro-typography ----------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[final]{microtype}

% ---------- Page geometry ----------
\usepackage[letterpaper,margin=1in]{geometry}

% ---------- Math & symbols ----------
\usepackage{amsmath,amssymb,mathtools}

% ---------- Colors, graphics, layout ----------
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage[most]{tcolorbox}
\tcbuselibrary{breakable}
\tcbset{breakable}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning}
\usetikzlibrary{calc,shadows.blur,decorations.pathmorphing}

% ---------- Tables ----------
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{longtable}

% ---------- Lists ----------
\usepackage{enumitem}
\setlist{topsep=4pt,itemsep=2pt,parsep=0pt,partopsep=0pt}
\setlist[description]{style=nextline,leftmargin=2.2em,labelsep=0.6em}

% ---------- Hyperlinks & references ----------
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=[rgb]{0.10,0.25,0.55},
  citecolor=[rgb]{0.10,0.45,0.25},
  urlcolor=[rgb]{0.55,0.10,0.10},
  pdfauthor={Fate's Edge},
  pdftitle={Fate's Edge --- Travel Guide and World Book},
  pdfcreator={LaTeX}
}
\usepackage{bookmark}
\usepackage[nameinlink,noabbrev,capitalise]{cleveref}

% ---------- Indexing ----------
\usepackage{imakeidx}
\makeindex

% Inline "print + index" helper
\newcommand{\indexterm}[2][]{%
  \textit{#2}%
  \if\relax\detokenize{#1}\relax\index{#2}\else\index{#1}\fi
}

% ---------- Table helpers & house style ----------
\renewcommand{\arraystretch}{1.15}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\definecolor{FeRow}{HTML}{F5F7FA}
\newcommand{\feTableStart}{\rowcolors{2}{FeRow}{white}\small}
\newcommand{\feTableEnd}{\rowcolors{2}{}{}}

% ---------- Small conveniences ----------
\newcommand{\textrarr}{\ensuremath{\rightarrow}}
\newcommand{\textlarr}{\ensuremath{\leftarrow}}
\newcommand{\textdarr}{\ensuremath{\downarrow}}
\newcommand{\beat}{\ensuremath{\triangleright}\,}
\newcommand{\fullsquare}{\raisebox{0.2ex}{\rule{0.9ex}{0.9ex}}}
\newcommand{\Tag}[1]{\texttt{[#1]}}

% ============================================================
% Branding & Metadata (single source of truth)
% ============================================================
\usepackage{etoolbox} % for \ifdefempty, \ifdefstring

% Brand palette (define once)
\definecolor{fePrimary}{HTML}{0D1B2A}   % deep blue-black
\definecolor{feAccent}{HTML}{6FFFE9}    % teal accent
\definecolor{feInk}{HTML}{101418}       % near-black
\definecolor{fePaper}{HTML}{F7F7F7}     % light paper

% Metadata variables + setters
\newcommand\feTitle{Fate's Edge}
\newcommand\feSubtitle{Comprehensive Resource Guide}
\newcommand\feTagline{}
\newcommand\feVersion{}
\newcommand\fePublisher{}
\newcommand\feCoverArt{} % optional path

\newcommand{\feSetTitle}[1]{\renewcommand\feTitle{#1}}
\newcommand{\feSetSubtitle}[1]{\renewcommand\feSubtitle{#1}}
\newcommand{\feSetTagline}[1]{\renewcommand\feTagline{#1}}
\newcommand{\feSetVersion}[1]{\renewcommand\feVersion and World Book{#1}}
\newcommand{\feSetPublisher}[1]{\renewcommand\fePublisher{#1}}
\newcommand{\feSetCoverArt}[1]{\renewcommand\feCoverArt{#1}}

% Text shadow helper (4 args: dx, dy, color, text)
\newcommand{\textshadow}[4]{%
  \begingroup
    \rlap{\raisebox{#2}{\hspace*{#1}\textcolor{#3}{#4}}}#4%
  \endgroup
}
% Light small-caps tracking
\newcommand{\FEsc}[1]{\textls[30]{\textsc{#1}}}

% =========================
% Basic Title Page
% =========================
\newcommand{\feTitlePage}{%
  \begin{titlepage}
    \centering
    \vspace*{3cm}
    
    {\Huge\bfseries Fate's Edge \par}
    \vspace{1cm}
    {\LARGE Comprehensive Resource Guide \par}
    \vspace{0.5cm}
    {\large \itshape Narrative-First Roleplaying System \par}
    
    \vfill
    
    {\large Version 1.0 \par}
    \vspace{0.2cm}
    {\large \today \par}
    \vspace{1cm}
    {\large \textsc{Nicholas A. Gasper}\par}
    
  \end{titlepage}
}

\begin{document}
% booktabs, tabularx, xcolor, tcolorbox, enumitem, hyperref, amsmath
\chapter{Travel Reference}

\section{Core Travel Procedure}

For each leg of a journey, draw 3--4 cards using the decks for your destination and controlling authority.

\begin{itemize}
\item Spade from the destination deck: sets the scene (place).
\item Heart from the destination deck: introduces the local actor or faction.
\item Club from the Wilds (general hazards) or destination (if strongly policed): brings pressure.
\item Diamond from the authority that gates the route: papers, escorts, rights, or exceptions.
\end{itemize}

Set a travel timer by the highest rank:
\begin{itemize}
\item 2--5 $\Rightarrow$ 4 segments
\item 6--10 $\Rightarrow$ 6 segments
\item J/Q/K $\Rightarrow$ 8 segments
\item A $\Rightarrow$ 10 segments
\end{itemize}

On success, advance to the next leg; on failure, mark delay, debt, or diversion and resolve a consequence in the fiction.

\section{Mode Rules}

\subsection{Sea Legs}
(Amaranthine/Dolmis/Aberderrin). If Theona or Valewood 9s show up anywhere in the seed, add an omission or taboo to the leg (a missing step, an unsaid name).

\subsection{Passes Underways}
(Aeler). Any A may convert a surface route to an under-route. Diamond from Aeler always codifies outcomes---no roll---for sealed doors, breath time, and escorted segments.

\subsection{Rivers}
Bridges, booms, and law in Ecktoria/Viterra; reed-mazes and bell-lines in Mistlands/Linn waters.

\subsection{Frontier Blends}
When origin and destination disagree on law, draw two Diamonds (one from each law) and choose which you will be judged by at the end of the leg.

\section{Regional Routes}

\subsection{Amaranthine Coastway}
Kahfagia $\rightarrow$ Ecktoria $\rightarrow$ Acasia $\rightarrow$ Marcott (Vhasia) $\rightarrow$ Fairport (Viterra).

\begin{itemize}
\item Draw Spade+Heart from the port you are entering.
\item Draw Club from the Wilds (gales, quarantine, harbor booms).
\item Draw Diamond from the port's deck (permits, berths), or Kahfagia if convoy/lantern law applies.
\end{itemize}

Note: Silkstrand is Acasia's sole metropolis and a major trading hub; even land caravans risk the marches to reach it.

\subsection{Astroegro Straits}
Thepyrgos controls the hinge between seas.

\begin{itemize}
\item Draw Spade+Heart and Club from Thepyrgos (stairs, boom, synod).
\item Draw Diamond from Thepyrgos (harbor pass, crane allotment).
\item Switch to Dolmis/Amaranthine modules once the boom is lowered.
\end{itemize}

\subsection{Dolmis Circuits}
Fairport (Viterra) $\rightarrow$ Theona (Three Greens) $\rightarrow$ Ubral fjords $\rightarrow$ Aelinnel west shore $\rightarrow$ Zakov (north coast).

\begin{itemize}
\item Spade+Heart from destination; Club from Linn or the Wilds; Diamond from Theona when island custom matters.
\item Zakov: Draw Spade+Heart from Zakov deck; Club from Zakov or Linn; Diamond from Zakov (harbor-green, corsair charter).
\end{itemize}

\subsection{Aelerian Passes Underways}
Vhasia/Viterra/Ubral south slopes $\rightarrow$ Aeler gates $\rightarrow$ Mistlands.

\begin{itemize}
\item Spade+Heart and Club from Aeler; Diamond from Aeler (Underway pass, Key-Writ, Breath time).
\end{itemize}

\subsection{Shadow Corridors}
Thin Shore (Valewood east coast): risky misted corridor north--south toward Zakov.

\begin{itemize}
\item Spade from Valewood, Heart from Mistlands (or Valewood), Club from Mistlands (wraiths, bell-line failures), Diamond from either Mistlands (Ward-salt, Lantern Writ) or Valewood (Way-cord, Truce-bough) depending on which law you invoke.
\item Rule of 9s: any 9 in the seed adds an omission.
\end{itemize}

Green Gate hops (Aelinnel): tide-timed arches.

\begin{itemize}
\item Spade+Club from Aelinnel (tides, geasa), Diamond from Aelinnel (Tide-path Key, Host-rights).
\end{itemize}

\subsection{River Roads}
Belworth: forms the boundary between Vhasia and Viterra. Twin ports at the mouth (Marcott and Fairport); at mid-reach Tarlington Falls requires lift/portage up to the plateau town. The Lence joins here from Vhasia. Use Viterra for east-bank audits/bridges, Vhasia for west-bank politics.

Yloka: raids run down from the north into the Aberderrin and Dolmis---overlay Linn Clubs and draw destination for Spade+Heart.

Acasian rivers: hazardous but profitable approaches to Silkstrand; use Acasia Clubs (feuds/levies/curse) and Diamonds (condotta, bridge rights).

\subsection{Steppe Frontiers}
Ykrul $\leftrightarrow$ Vilikari $\leftrightarrow$ Ecktoria/Acasia borders $\leftrightarrow$ Black Banner territories.

\begin{itemize}
\item Spade from the road you ride (Wolf Road or Foedus Stone), Heart from the counterpart (envoys, comitatus), Club from Wilds or frontier decks (rasputitsa, hostage protocol, foedus recall), Diamond from either Ykrul (paiza, remounts) or Vilikari (foedus seal, bench-right).
\item Black Banner: Draw Spade from Ykrul/Vilikari decks (shared territory), Heart from Black Banner, Club from Wilds or frontier decks, Diamond from Ykrul/Vilikari/Black Banner.
\end{itemize}

\section{Gateways Control Points}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
\textbf{Gate} & \textbf{Deck} & \textbf{Diamond Source} & \textbf{Typical Leverage} \\
\midrule
Kassamira (Kahfagia) & Kahfagia & Kahfagia & Convoy letter; lantern-law warrant; pilotage token \\
Ecktoria Capital Port & Ecktoria & Ecktoria & Berth priority; coin-house credit; procession permit \\
Silkstrand & Acasia & Acasia & Exchange floor; condotta; watergate priority \\
Marcott & Vhasia & Vhasia & Letters patent; safeconduct; fairground precedence \\
Fairport & Viterra & Viterra & Customs seal; ferry priority; Dawn escort \\
Tarlington (plateau lift) Falls & Viterra & Viterra & Lift priority; bridge/portage pass; County Thing ruling \\
Thepyrgos (boom) & Thepyrgos & Thepyrgos & Harbor pass; crane allotment; Synod indulgence \\
High-Mist Pass & Aeler & Aeler & Underway pass; Key-Writ; Breath time allotment \\
Payden's Port & Mistlands & Mistlands & Lantern writ; ward-salt; wraith indemnity \\
Thin Shore (Valewood) & Valewood & Valewood/Mistlands & Way-cord; dew-mirror; lantern writ \\
Green Gate (Aelinnel) & Aelinnel & Aelinnel & Tide-path key; host-right bough; salt license \\
Ubral Passes & Ubral & Aeler/Ubral & Vurim pass-ring; hill-fort shelter writ; guide's braid \\
Thing-holm (Linn) & Linn & Linn & Thing ruling; harbor-green; escort writ \\
Foedus Stone (March) & Vilikari & Vilikari/Ykrul & Foedus renewal; remount allotment; hostage treaty \\
Zakov Roadstead & Zakov & Zakov & Harbor-green chit; corsair charter; pilot token \\
Black Banner Camps & Black Banner & Black Banner/Ykrul/Vilikari & Safe-conduct; remount chit; clan truce \\
\bottomrule
\end{tabular}
\caption{Gateway Control Points}
\end{table}

% =======================
% ZAKOV OVERLAY + ITINERARY
% =======================

\section{Overlay: Gatekeepers (Zakov Roadstead \& Harbour Quarters)}
Draw \textbf{one} micro-card each time you \emph{enter}, \emph{shift wharfs}, or \emph{depart} Zakov. Combine with the normal seed.

\begin{tcolorbox}[enhanced,breakable,sharp corners,boxrule=.6pt,title={Zakov Gatekeepers Micro-Cards (draw 1)}]
\begin{enumerate}[label=\arabic*.]
  \item \textbf{Bureaucrat Quirk} --- stamps only at \emph{eighth bell}; night filings need a \emph{Night Magistrate} countersign.
  \item \textbf{Required Stamp} --- \emph{Harbor-Green} chit for boom passage this tide; without it, DV +1 on all port checks.
  \item \textbf{Preferred Gratuity} --- figs or lamp oil; wrong gift offends, marking 1 \emph{Exposure} on your port reputation.
  \item \textbf{Audit Trap} --- ``contraband tallies''; if carrying \emph{bell-metal} or \emph{dyestuffs}, start a 4-seg \emph{Inspection} timer.
  \item \textbf{Local Courtesy} --- speak the \emph{Dock Oath} (short pledge not to knife on planks) or suffer \emph{Desperate} position in brawls.
  \item \textbf{Shortcut Rumor} --- \emph{Smuggler's Ladder} on the sea wall; spend a \SuitDiamond{} (\emph{Ladder Map}) to skip one queue.
  \item \textbf{Pilot's Privilege} --- a \emph{Pilot Token} bumps you ahead on boom lift; otherwise pay tow or lose 1 segment.
  \item \textbf{Silent Court} --- the \emph{Black Bishop} is sitting tonight; legal Clubs escalate by +1 SB if you contest a ruling.
\end{enumerate}
\end{tcolorbox}

\section{Special Rules \& Taboos}

\subsection{Theona Valewood 9s}
Whenever a 9 appears in a seed involving Theona or Valewood, add an omission: a missing step, an unsaid name, an unseen guest. If the 9 is a Diamond, you may break the taboo once---someone will come to collect.

\subsection{Aeler Aces}
An A can rewire routes beneath the mountains. You may swap your Spade to Aeler (keeping the other seed cards) and continue under-vault procedures.

\subsection{Echoing Omens}
Any A adds a motif you can echo later (ash on the wind, a bell heard underground, a landmark that returns elsewhere). Make the world feel like it remembers.

\section{Worked Itineraries}

\subsection{West-to-East Coastal Haul (Kahfagia $\rightarrow$ Viterra)}

\textbf{Leg 1: Kassamira $\rightarrow$ Ecktoria}

Seed: Ecktoria Spade (Aqueduct arcades), Ecktoria Heart (Coin-house factor), Wilds Club (gale), Kahfagia Diamond (convoy letter).

Clock: 6. Read: cross under coin-house eyes with convoy priority; bad weather strains schedules.

\textbf{Leg 2: Ecktoria $\rightarrow$ Silkstrand}

Seed: Acasia Spade (Three-Queens Bridge), Acasia Heart (Dyers' Guildmistress), Acasia Club (loom strike), Acasia Diamond (Exchange pass).

Clock: 6--8 depending on unrest.

\textbf{Leg 3: Silkstrand $\rightarrow$ Marcott}

Seed: Vhasia Spade (Pont-du-Tithe), Vhasia Heart (Parlement clerk), Vhasia Club (coin rumor), Vhasia Diamond (letters patent).

\textbf{Leg 4: Marcott $\rightarrow$ Fairport}

Seed: Viterra Spade (Fairport tideworks), Viterra Heart (shipwright), Linn Club (boom lifts), Viterra Diamond (customs seal).

Clock: 6.

\subsection{Acasia $\rightarrow$ Mistlands (Forgotten Pass + Under-Gate)}

\textbf{Leg 1: Silkstrand $\rightarrow$ Aeler Gate}

Seed: Aeler Spade (Avalanche gallery), Aeler Heart (Geometer), Aeler Club (Engineer requisition), Aeler Diamond (Underway Pass).

\textbf{Leg 2: Gate $\rightarrow$ Mistlands}

Seed: Mistlands Spade (Bell-Line levee), Mistlands Heart (Bell-warden), Mistlands Club (wraith crossing), Mistlands Diamond (Ward-salt).

\subsection{Thin Shore $\rightarrow$ Zakov Corridor}

\textbf{Leg 1: Payden's Port $\rightarrow$ Thin Shore}

Seed: Valewood Spade (Green lane), Mistlands Heart (Protectorate clerk), Mistlands Club (bell failure), Mistlands Diamond (Lantern writ).

\textbf{Leg 2: Thin Shore transit}

Seed: Valewood Spade (Unfound stile), Valewood Heart (Path-warden), Valewood Club (Sweet wind), Valewood Diamond (Way-cord).

Rule of 9s applies.

\subsection{Frontier Mediation (Ykrul $\leftrightarrow$ Vilikari)}

\textbf{Leg: Foedus Stone parley}

Seed: Vilikari Spade (Foedus Stone), Ykrul Heart (Noyan envoy), Wilds Club (rasputitsa), Vilikari Diamond (Foedus Renewal).

Blend: also draw a Ykrul Diamond (Paiza). Choose which law saves your schedule.

\subsection{Worked Itinerary: Thin Shore $\rightarrow$ Zakov $\rightarrow$ Theona (Corsair Jobs)}
A fast arc for crews running the misted coast into pirate politics and back into isle taboos.

\textbf{Leg 1: Payden's Port $\rightarrow$ Thin Shore (Shadow Corridor)}
\begin{Leg}{Seed}
Valewood \SuitSpade{} (Green lane / Unfound stile);
Mistlands \SuitHeart{} (Protectorate clerk);
Mistlands \SuitClub{} (bell-line failure);
Mistlands \SuitDiamond{} (Lantern Writ).
\end{Leg}
\textbf{Clock:} 6. \textbf{Rule of 9s:} any 9 adds an omission (missing step/name).

\textbf{Leg 2: Thin Shore Transit (Skirting toward Zakov)}
\begin{Leg}{Seed}
Valewood \SuitSpade{} (Sea-mist arcade);
Valewood \SuitHeart{} (Path-warden);
Valewood \SuitClub{} (Sweet wind);
Valewood \SuitDiamond{} (Way-cord).
\end{Leg}
\textbf{Clock:} 6. \textbf{Note:} spending \emph{Way-cord} negates one \emph{Sweet wind} lie.

\textbf{Leg 3: Approach to Zakov (Roadstead \& Booms)}
\begin{Leg}{Seed}
Zakov \SuitSpade{} (Boomhouse or Red Wharf);
Zakov \SuitHeart{} (Pilot-Matron \emph{or} Night Magistrate);
Zakov \SuitClub{} (Boom Drop \emph{or} Customs Sweep);
Zakov \SuitDiamond{} (Harbor-Green Chit \emph{or} Pilot Token).
\end{Leg}
\textbf{Clock:} 6--8 (apply the Gatekeepers overlay draw on arrival). \textbf{If a 9 shows:} \emph{Missing Ninth}---remove one critical step/name until paid in favors.

\textbf{Leg 4: Corsair Job Inside Zakov (Pick 1)}
\begin{Leg}{A) Lift a Hull from Drydock Four}
Zakov \SuitSpade{} (Drydock Four);
Zakov \SuitHeart{} (Corsair Quartermaster);
Zakov \SuitClub{} (Admiralty Audit);
Zakov \SuitDiamond{} (Shipwright's Lien Release).
\end{Leg}
\textbf{Goal:} tow a seized hull at night; \emph{Lien Release} avoids bond but flips a debt later.

\begin{Leg}{B) Court the Black Bishop for Indulgence}
Zakov \SuitSpade{} (Black Bell Tower);
Zakov \SuitHeart{} (Black Bishop);
Zakov \SuitClub{} (Bounty Proclamation);
Zakov \SuitDiamond{} (Magistrate's Hush).
\end{Leg}
\textbf{Goal:} seal a case/docket so a rival can't use it. \textbf{Hook:} rumor must be paid (coin or gossip).

\begin{Leg}{C) Smuggler's Ladder Run}
Zakov \SuitSpade{} (Lantern Ladder);
Zakov \SuitHeart{} (Lampman);
Zakov \SuitClub{} (Informant Flip);
Zakov \SuitDiamond{} (Smuggler's Ladder Map).
\end{Leg}
\textbf{Goal:} bypass boom eyes with a wall-stair. \textbf{Complication:} your tip was sold twice.

\textbf{Leg 5: Zakov $\rightarrow$ Theona (Isles \& Moot)}
\begin{Leg}{Seed}
Theona \SuitSpade{} (Uncounted Bridge);
Theona \SuitHeart{} (Matron of Wells \emph{or} Moot Envoy);
Linn \SuitClub{} (fogfall raids);
Theona \SuitDiamond{} (Moot Token).
\end{Leg}
\textbf{Clock:} 6--8. \textbf{Taboo:} do \emph{not} count the steps aloud; breaking it imposes \emph{Controlled} on all isle bargains this scene.

\textbf{Leg 6: Theona Contract (Pick 1)}
\begin{Leg}{A) Raid-Truce at the Skerries}
Theona \SuitSpade{} (Tide caves);
Theona \SuitHeart{} (Isle Moot Envoy);
Theona \SuitClub{} (Muster drum);
Theona \SuitDiamond{} (Raid-truce Ribbon).
\end{Leg}
\textbf{Goal:} secure safe passage window; failure triggers a Linn muster timer.

\begin{Leg}{B) Deliver the Ledger Shard}
Theona \SuitSpade{} (Well-yard);
Theona \SuitHeart{} (Matron of Wells);
Zakov \SuitClub{} (Debt Call follows you);
Theona \SuitDiamond{} (Sanctuary Night).
\end{Leg}
\textbf{Goal:} hide a Zakov debt long enough to copy data; sanctuary buys time.

\subsection{Steppe Passage: Black Banner Territory}
A dangerous journey through contested lands where three powers vie for control.

\textbf{Leg: Foedus Stone $\rightarrow$ Black Banner Camp}
\begin{Leg}{Seed}
Ykrul/Vilikari \SuitSpade{} (Wolf Road milepost or Foedus Stone);
Black Banner \SuitHeart{} (Clan Elder or War Captain);
Wilds \SuitClub{} (Rasputitsa or Remount Sickness);
Black Banner \SuitDiamond{} (Safe-conduct or Remount Chit).
\end{Leg}
\textbf{Clock:} 6--8. \textbf{Complication:} Foedus recall may invalidate your papers.

\textbf{Leg: Black Banner Camp $\rightarrow$ Ykrul Territory}
\begin{Leg}{Seed}
Ykrul \SuitSpade{} (Winter camp ring or Khagan's way-station);
Ykrul \SuitHeart{} (Khatun of the Ring or Noyan envoy);
Frontier \SuitClub{} (Hostage protocol or Feud spark);
Ykrul \SuitDiamond{} (Paiza tablet or Foedus seal).
\end{Leg}
\textbf{Clock:} 6--8. \textbf{Note:} Choose which law applies---Black Banner customs or Ykrul traditions.

\section{Pace, Timers, and Consequences}

\begin{itemize}
\item Road/river leg: 6-segment timer; +2 segments if crossing a law boundary (new passports, new prayers).
\item Sea strait/mountain pass: 8-segment timer; --2 segments if you spend a Diamond that codifies priority (pilotage, pass, convoy).
\item Shadow corridor: 6-segment timer; on any A, add a free omen to echo later.
\item Failure defaults: delay (lose time; an enemy advances), diversion (alternate route; new Club), or debt (promise, fine, or favor owed to the authority that controls the gate).
\end{itemize}

\noindent\textbf{SB Note.} Re-rolling 1s does \emph{not} erase their SB; any new 1s on the re-roll add more SB.

\section{Story Beat (SB) Menus}

\subsection{Universal SB Spend Menu}
\begin{itemize}
  \item \textbf{1 SB} --- Shave time/position: add +1 segment to the current timer \emph{or} drop Position one step \emph{or} impose a small fee/toll \emph{or} mark 1 step of \emph{Asset Exposure}.
  \item \textbf{2 SB} --- Escalate the pressure: upgrade the active \SuitClub{} (e.g., ``inspection'' $\rightarrow$ ``seizure attempt'') \emph{or} force a \emph{choice of Diamonds} (pick which law you'll be judged by) \emph{or} tick a relevant \emph{Front} by 1.
  \item \textbf{3 SB} --- Lock a gate/taboo: close a route (boom drops; pass closes) until a named \SuitDiamond{} is spent \emph{or} invoke a regional taboo (Theona/Valewood 9s; oath backlash) with teeth \emph{or} auto-trigger a \emph{Debt Flip} on a Diamond already in play.
  \item \textbf{4+ SB} --- Reshape the leg: rewrite the route (Aeler under-way reroute; sea lane reversal), start a \emph{new} 4--6 segment complication timer \emph{and} tick a \emph{Front} by +1; demand immediate payment on a Patron/Asset Exposure (lose it if refused).
\end{itemize}
\textit{End-of-leg conversion (optional):} Unspent SB convert to \emph{world motion} at a rate of \textbf{3 SB $\rightarrow$ tick 1 Front} (GM choice).

\subsection{Sea (Amaranthine/Dolmis/Aberderrin)}
\begin{itemize}
  \item \textbf{1 SB:} Fogfall nips visibility; DV +1 for the next pilotage check.
  \item \textbf{2 SB:} Boom shifts; pilot required (spend a \emph{pilotage} \SuitDiamond{} or delay +2 segments).
  \item \textbf{3 SB:} Lantern-law change mid-leg; choose: pay fine \emph{or} reroute timer +2 \emph{or} spend \emph{Convoy Letter}.
  \item \textbf{4 SB:} Squall wall: split the convoy; lose an escort unless \emph{Moot token} or \emph{Escort} Diamond is spent.
\end{itemize}

\subsection{Passes \& Underways (Aeler)}
\begin{itemize}
  \item \textbf{1 SB:} White wind; breath-time check (mark Fatigue or slow +1 segment).
  \item \textbf{2 SB:} Engineer requisition; surrender cargo slot \emph{or} coin \emph{or} DV +1 until the gate.
  \item \textbf{3 SB:} Under-audit; must go below (swap Spade to Aeler) unless you spend \emph{Key-Writ}.
  \item \textbf{4 SB:} Cave-in detour adds a fresh 4-segment timer unless \emph{Underway Pass} is spent.
\end{itemize}

\subsection{Rivers (Belworth \& friends)}
\begin{itemize}
  \item \textbf{1 SB:} Sandbar drift; portage a beat or pay for a tow.
  \item \textbf{2 SB:} Lift strike at Tarlington; need \emph{Bridge/Portage Pass} or delay +2 segments.
  \item \textbf{3 SB:} Cross-mouth embargo (Marcott/Fairport); choose which harbor's law applies (draw/commit its \SuitDiamond{}).
  \item \textbf{4 SB:} Ice-break or flood surge; split cargo (random item at risk) unless an \emph{Escort} Diamond is spent.
\end{itemize}

\subsection{Shadow Corridors (Valewood/Mistlands)}
\begin{itemize}
  \item \textbf{1 SB:} Bell-line stutter; next guidance roll at Disadvantage (or DV +1).
  \item \textbf{2 SB:} Omission manifests; something necessary is ``missing'' until \emph{Ward-salt} or \emph{Way-cord} is spent.
  \item \textbf{3 SB:} Host-law claims you as \emph{host/guest}; accept a service timer or pay with a memory (lose a minor Contact until end of leg).
  \item \textbf{4 SB:} Wraith procession crosses your path; create a 6-segment \emph{Don't Look Back} timer; spending \emph{Lantern Writ} downgrades it to 4.
\end{itemize}

\subsection{Steppe Frontiers (Ykrul/Vilikari/Black Banner)}
\begin{itemize}
  \item \textbf{1 SB:} Remount fatigue; lose 1 ``fresh mount'' use or slow +1 segment.
  \item \textbf{2 SB:} Hostage protocol; must place/accept a token-kin until the next gate (or spend \emph{Safe-hostage}).
  \item \textbf{3 SB:} Foedus recall; your papers conflict---pick Ykrul \emph{Paiza} or Vilikari \emph{Foedus Seal} and offend the other.
  \item \textbf{4 SB:} Muster checkpoint; disarm unless traveling under \emph{Standard Protection} or \emph{Khagan's Writ}.
  \item \textbf{Black Banner Specific:} Clan feud erupts; choose sides or pay double tolls.
\end{itemize}

\subsection{Ports \& Forums (Marcott, Fairport, Silkstrand, Thepyrgos, Zakov)}
\begin{itemize}
  \item \textbf{1 SB:} Clerk error; pay a ``stamp'' fee or answer questions (DV +1).
  \item \textbf{2 SB:} Rival petition; they jump the queue unless you spend a Warrant/Letters \SuitDiamond{}.
  \item \textbf{3 SB:} Audit cascade; open a new 4-segment \emph{Inspection} timer tied to cargo/ledger.
  \item \textbf{4 SB:} Edict flare; convert today's \SuitClub{} into a temporary ``law'' everyone must obey or take on a \emph{Debt Flip}.
  \item \textbf{Zakov Specific:} Silent Court ruling; legal complications escalate by +1 SB.
\end{itemize}

\subsection{Diamond \& Exposure Interactions}
\begin{itemize}
  \item \textbf{2 SB:} Flip any \SuitDiamond{} spent this leg: its \emph{Debt Flip} is collected now (not later).
  \item \textbf{2 SB:} Mark 1 step on a named Asset/Patron's \emph{Exposure} track.
  \item \textbf{3 SB:} Downgrade an \emph{Intricate} success to \emph{Detailed} (remove the flourish or DV −1 benefit).
  \item \textbf{4 SB:} Seize or suspend a visible Diamond's benefit until another is played (e.g., convoy letter overridden by emergency edict).
\end{itemize}
\clearpage
\section{Location Decks for Navigation}
% ============================================================
% Travel Reference Chapter (to be inserted after "On Obligation and Trade")
% ============================================================

\newpage
\chapter{Travel Reference}
\label{chap:travel-reference}

\begin{quote}
\small
\textit{``The road is not a ledger. It is a conversation. Speak carefully. Listen closely. And always leave something behind for the next traveler.''} \\
--- Dusana of the Raven Road
\end{quote}

\section{Core Travel Procedure}

For each leg of a journey, draw 3--4 cards using the decks for your destination and controlling authority.

\begin{itemize}
\item \spadesuit{} from the destination deck: sets the scene (place).
\item \heartsuit{} from the destination deck: introduces the local actor or faction.
\item \clubsuit{} from the Wilds (general hazards) or destination (if strongly policed): brings pressure.
\item \diamondsuit{} from the authority that gates the route: papers, escorts, rights, or exceptions.
\end{itemize}

Set a travel timer by the highest rank:
\begin{itemize}
\item 2--5 $\Rightarrow$ 4 segments
\item 6--10 $\Rightarrow$ 6 segments
\item J/Q/K $\Rightarrow$ 8 segments
\item A $\Rightarrow$ 10 segments
\end{itemize}

On success, advance to the next leg; on failure, mark delay, debt, or diversion and resolve a consequence in the fiction.

\section{The Amaranthine: A Textual Map}

The Amaranthine is isomorphic to Earth's geography:

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Earth} & \textbf{Amaranthine} \\
\hline
Northern Europe & Violet Steppe \\
The Alps & Aelerian Mountains \\
Iberia + Morocco & Kahfagia \\
Italy + France + Germany & Ecktoria \\
Holy Roman Empire & Acasia \\
Western Balkans & Vhasia \\
Eastern Balkans & Viterra (straddles Dolmis \& Amaranthine) \\
Southwest Balkans & Thepyrgos \\
Black Sea & The Dolmis \\
Caspian Sea & The Aberderrin \\
North Shore (Black Sea) & Zakov \\
East Shore (Black Sea) & Aelinnel, Aelaerem \\
South Shore (Black Sea) & Viterra, Thepyrgos \\
North Shore (Caspian) & The Linns \\
West Shore (Caspian) & Mistlands (Payden's Port) \\
South of Mistlands & Aeler \\
Between Mistlands \& Steppe & Direwood \\
North of Aelaerem/Aelinnel, East of Aberderrin & Valewood \\
Coastal strip where Valewood pulled back & Thin Coast \\
North of Viterra, South of Aelerian Mountains & Ubral \\
\hline
\end{tabular}
\end{center}

\textbf{Wind Patterns:} The Violet Steppe's winds are \emph{reversed} from our world. This affects trade routes across the steppe, sea routes in the Aberderrin, and the Direwood's weather (storms come from the east).

\subsection{Textual Map}

\begin{verbatim}
                     ┌─────────────────────────────────────────────────────────┐
                     │                  THE VIOLET STEPPE                     │
                     │         (Northern Europe, reversed winds)              │
                     │                                                         │
                     │    ┌──────────────┐         ┌──────────────┐           │
                     │    │   YKRUL      │         │   KUVANI     │           │
                     │    │ (Wolf Banners)│         │ (Remounts)   │           │
                     │    └──────┬───────┘         └──────┬───────┘           │
                     │           │                        │                     │
                     │           │    DIREWOOD           │                     │
                     │           │   (Folk Horror)        │                     │
                     │           │         │             │                     │
                     └───────────┼─────────┼─────────────┼─────────────────────┘
                                 │         │             │
                ┌────────────────┼─────────┼─────────────┼────────────────────┐
                │                │         │             │                    │
                │    MISTLANDS   │         │   AELER     │                    │
                │  (Bells, Salt) │         │  (Dwarves)  │     VALEWOOD       │
                │    Payden's    │         │             │   (Empire Under    │
                │      Port      │         │             │      Leaves)       │
                │         │      │         │             │       │            │
                │         │      └─────────┼─────────────┼───────┼────────────┤
                │         │                │             │       │            │
                │         │    AELERIAN    │   UBRAL     │ THIN  │            │
                │         │   MOUNTAINS    │ (Highland   │ COAST │            │
                │         │   (The Alps)   │   Clans)    │       │            │
                │         │                │     │       │       │            │
                │         └────────────────┼─────┼───────┼───────┼────────────┤
                │                          │     │       │       │            │
                │        ┌─────────────────┼─────┼───────┼───────┼────────────┤
                │        │                 │     │       │       │            │
                │        │    ACASIA       │VHASIA│VITERRA│AELINNEL│ AELAEREM │
                │        │  (Broken       │(West │(East  │(Gnomes)│(Halflings)│
                │        │   Marches)     │Balkans)│Balkans)│        │          │
                │        │      │         │  │   │   │   │        │          │
                │        │      │         │  │   │   │   │        │          │
                │        │ SILKSTRAND     │  │   │   │   │        │          │
                │        │  (City of      │  │   │   │   │        │          │
                │        │  Bridges)      │  │   │   │   │        │          │
                │        │      │         │  │   │   │   │        │          │
                │        └──────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │               │         │  │   │   │   │        │          │
                │    ┌──────────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │    │          │         │  │   │   │   │        │          │
                │    │ ECKTORIA │         │  │   │   │   │        │          │
                │    │ (Marble &│         │  │   │   │   │        │          │
                │    │   Fire)  │         │  │   │   │   │        │          │
                │    │          │         │  │   │   │   │        │          │
                │    └──────────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │               │         │  │   │   │   │        │          │
                │    ┌──────────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │    │ KAHFAGIA │         │  │   │   │   │        │          │
                │    │(Creole   │         │  │   │   │   │        │          │
                │    │Thalassoc.)│         │  │   │   │   │        │          │
                │    │          │         │  │   │   │   │        │          │
                │    └──────────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │               │         │  │   │   │   │        │          │
                │    ┌──────────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │    │   THE     │         │  │   │   │   │        │          │
                │    │ AMARAN-   │         │  │   │   │   │        │          │
                │    │  THINE    │         │  │   │   │   │        │          │
                │    │   SEA     │         │  │   │   │   │        │          │
                │    │ (Mediter-│         │  │   │   │   │        │          │
                │    │  ranean)  │         │  │   │   │   │        │          │
                │    │    │      │         │  │   │   │   │        │          │
                │    │    │      │         │  │   │   │   │        │          │
                │    └──────┼────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │           │    │         │  │   │   │   │        │          │
                │    ┌──────┼────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │    │      │    │THEPYRGOS│  │   │   │   │        │          │
                │    │      │    │(City of │  │   │   │   │        │          │
                │    │      │    │ Stairs) │  │   │   │   │        │          │
                │    │      │    │         │  │   │   │   │        │          │
                │    └──────┼────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │           │    │         │  │   │   │   │        │          │
                │    ┌──────┼────┼─────────┼──┼───┼───┼───┼────────┼──────────┤
                │    │      │    │ THEONA  │  │   │   │   │        │          │
                │    │ ZAKOV│    │ (Island)│  │   │   │   │        │          │
                │    │(North│    │         │  │   │   │   │        │          │
                │    │ Shore)│   └─────────┼──┼───┼───┼───┼────────┼──────────┤
                │    │      │              │  │   │   │   │        │          │
                │    │      │    THE       │  │   │   │   │        │          │
                │    │      │   DOLMIS SEA │  │   │   │   │        │          │
                │    │      │  (Black Sea) │  │   │   │   │        │          │
                │    │      │              │  │   │   │   │        │          │
                │    └──────┴──────────────┴──┴───┴───┴───┴────────┴──────────┘
                │
                │    ┌──────────┐
                │    │  LINNS   │
                │    │(Skerries)│
                │    │     │    │
                │    │     │    │
                │    │  ABERDERRIN SEA (Caspian Sea)
                │    │     │    │
                │    │     │    │
                │    └─────┴────┘
                │
                │    LINNS "SAIL OVER LAND" (Portage)
                │    around the mountains to reach the Dolmis
                │    via the Thin Coast
\end{verbatim}

\section{Route Tables}

\subsection{The Amaranthine Coastway (West to East)}

\textbf{Kahfagia $\rightarrow$ Ecktoria $\rightarrow$ Acasia $\rightarrow$ Silkstrand $\rightarrow$ Vhasia $\rightarrow$ Viterra $\rightarrow$ Thepyrgos}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Kassamira (Kahfagia) & Ecktoria Capital Port & Coastal shipping & Ecktoria & Convoy Letter & 6 \\
2 & Ecktoria & Silkstrand (Acasia) & Coastal shipping & Acasia & Exchange Pass & 6--8 \\
3 & Silkstrand & Marcott (Vhasia, west Balkans) & River/coastal & Vhasia & Letters Patent & 6 \\
4 & Marcott & Fairport (Viterra, east Balkans) & River/coastal & Viterra & Customs Seal & 6 \\
5 & Fairport & Thepyrgos (southwest corner) & Black River & Thepyrgos & Harbor Pass & 6--8 \\
\hline
\end{longtable}

\subsection{The Dolmis Circuit (Circum-Dolmis)}

\textbf{Fairport (Viterra) $\rightarrow$ Theona $\rightarrow$ Zakov $\rightarrow$ Aelinnel $\rightarrow$ Aelaerem $\rightarrow$ Thin Coast $\rightarrow$ Valewood $\rightarrow$ back to Fairport}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Fairport (Viterra, west shore) & Theona (island in Dolmis) & Island crossing & Theona & Moot Token & 6 \\
2 & Theona & Zakov (north shore) & Sea lane & Zakov & Harbor-Green & 6--8 \\
3 & Zakov & Aelinnel (east shore) & North coast & Aelinnel & Tide-Path Key & 6 \\
4 & Aelinnel & Aelaerem (south of Aelinnel) & Coast road & Aelaerem & Host-Right Bough & 4 \\
5 & Aelaerem & Thin Coast (Valewood coast) & Coastal strip & Valewood & Way-Cord & 6 \\
6 & Thin Coast & Valewood interior & Forest road & Valewood & Truce-Bough & 6 \\
\hline
\end{longtable}

\subsection{The Aelerian Passes (North-South Through the Mountains)}

\textbf{Vhasia/Viterra/Ubral $\rightarrow$ Aeler Gates $\rightarrow$ Mistlands $\rightarrow$ Direwood $\rightarrow$ Violet Steppe}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Marcott (Vhasia) & Aeler Gate & Mountain pass & Aeler & Underway Pass & 6--8 \\
2 & Fairport (Viterra) & Aeler Gate & Mountain pass & Aeler & Underway Pass & 6--8 \\
3 & Ubral (north of Viterra) & Aeler Gate & Mountain pass & Aeler & Pass-Ring & 6--8 \\
4 & Aeler Gate & Payden's Port (Mistlands) & Underway/crossing & Mistlands & Ward-Salt & 6 \\
5 & Payden's Port & Direwood (between Mistlands \& Steppe) & Forest crossing & Mistlands & Lantern Writ & 6--8 \\
6 & Direwood & Violet Steppe & Forest-to-steppe & Ykrul & Foedus Seal & 8 \\
\hline
\end{longtable}

\subsection{The Shadow Corridors (Valewood Coast)}

\textbf{Thin Coast $\leftrightarrow$ Valewood $\leftrightarrow$ Mistlands (north-south)}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Thin Coast (south) & Thin Coast (north) & Coastal strip & Valewood & Way-Cord & 6 \\
2 & Thin Coast (north) & Payden's Port (Mistlands) & Forest/shadow & Mistlands & Lantern Writ & 6--8 \\
3 & Thin Coast & Zakov & Sea/shadow & Zakov & Pilot Token & 6 \\
\hline
\end{longtable}

\subsection{The Linn Raid Routes (Portage)}

\textbf{Linn Skerries $\rightarrow$ Portage (over land) $\rightarrow$ Dolmis (via Thin Coast)}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Linn Skerries (Aberderrin north shore) & Thin Coast & Portage (sail over land) & Linn & Thing Ruling & 6 \\
2 & Thin Coast & Dolmis Sea & Sea lane & Zakov & Escort Writ & 6 \\
3 & Thin Coast & Zakov & Sea lane & Zakov & Pilot Token & 6 \\
4 & Thin Coast & Aelinnel (Dolmis east shore) & Sea lane & Aelinnel & Tide-Path Key & 6 \\
\hline
\end{longtable}

\subsection{The Steppe Frontiers}

\textbf{Ykrul $\leftrightarrow$ Vilikari $\leftrightarrow$ Ecktoria/Acasia $\leftrightarrow$ Black Banner Territories}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Ykrul Winter Camp & Foedus Stone (Vilikari) & Wolf Road & Vilikari & Foedus Renewal & 6--8 \\
2 & Foedus Stone & Ecktoria/Acasia border & Steppe road & Wilds & Remount Chit & 6 \\
3 & Ecktoria/Acasia border & Black Banner Camp & Frontier & Black Banner & Safe-Conduct & 6--8 \\
4 & Black Banner Camp & Ykrul Territory & Steppe road & Ykrul & Paiza Tablet & 6--8 \\
\hline
\end{longtable}

\subsection{The River Roads}

\textbf{Belworth:} Forms the boundary between Vhasia (west) and Viterra (east). The Black River forms most of the border between Viterra and Thepyrgos.

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Marcott (Vhasia) & Pont-du-Tithe & River & Vhasia & Letters Patent & 4--6 \\
2 & Pont-du-Tithe & Fairport (Viterra) & River & Viterra & Customs Seal & 4--6 \\
3 & Fairport & Black River bend & River & Viterra & Portage Pass & 6 \\
4 & Black River bend & Thepyrgos & River & Thepyrgos & Harbor Pass & 6 \\
\hline
\end{longtable}

\subsection{The Green Gate Hops (Aelinnel)}

\textbf{Aelinnel (east shore) $\leftrightarrow$ Aelaerem (south) $\leftrightarrow$ Valewood (north)}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Aelinnel Coast & Green Gate & Tide path & Aelinnel & Tide-Path Key & 4--6 \\
2 & Green Gate & Aelaerem Downs & Forest road & Aelaerem & Host-Right Bough & 4 \\
3 & Green Gate & Valewood & Phasing gate & Valewood & Way-Cord & 6 \\
\hline
\end{longtable}

\subsection{The Astroegro Straits (Thepyrgos Control)}

\textbf{Thepyrgos $\leftrightarrow$ Amaranthine $\leftrightarrow$ Dolmis (The hinge between seas)}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Thepyrgos & Astroegro Straits & Chain/boom & Thepyrgos & Harbor Pass & 6 \\
2 & Astroegro Straits & Amaranthine Sea & Sea lane & Kahfagia & Pilotage Token & 6--8 \\
3 & Astroegro Straits & Dolmis Sea & Sea lane & Zakov & Corsair Charter & 6--8 \\
\hline
\end{longtable}

\subsection{Ubral Routes (North of Viterra, South of Aelerian Mountains)}

\textbf{Ubral $\leftrightarrow$ Viterra $\leftrightarrow$ Aeler (via passes)}

\begin{longtable}{|p{1.5cm}|p{2.5cm}|p{2.5cm}|p{1.5cm}|p{2cm}|p{1.5cm}|}
\hline
\textbf{Leg} & \textbf{Origin} & \textbf{Destination} & \textbf{Gateway} & \textbf{Deck} & \textbf{Diamond} & \textbf{Timer} \\
\hline
1 & Ubral Highlands & Fairport (Viterra) & Mountain pass & Ubral & Guide's Braid & 4--6 \\
2 & Ubral Highlands & Aeler Gate & Mountain pass & Aeler & Pass-Ring & 6--8 \\
3 & Ubral Highlands & Vhasia (west) & Mountain pass & Ubral & Shelter Writ & 6 \\
\hline
\end{longtable}

\section{Gateway Control Points}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Gateway} & \textbf{Deck} & \textbf{Diamond Source} & \textbf{Typical Leverage} \\
\hline
Kassamira (Kahfagia) & Kahfagia & Kahfagia & Convoy letter; lantern-law warrant; pilotage token \\
Ecktoria Capital Port & Ecktoria & Ecktoria & Berth priority; coin-house credit; procession permit \\
Silkstrand & Acasia & Acasia & Exchange floor; condotta; watergate priority \\
Marcott (Vhasia) & Vhasia & Vhasia & Letters patent; safe-conduct; fairground precedence \\
Fairport (Viterra) & Viterra & Viterra & Customs seal; ferry priority; Dawn escort \\
Thepyrgos (boom) & Thepyrgos & Thepyrgos & Harbor pass; crane allotment; Synod indulgence \\
High-Mist Pass & Aeler & Aeler & Underway pass; Key-Writ; Breath time allotment \\
Payden's Port & Mistlands & Mistlands & Lantern writ; ward-salt; wraith indemnity \\
Thin Shore (Valewood) & Valewood & Valewood/Mistlands & Way-cord; dew-mirror; lantern writ \\
Green Gate (Aelinnel) & Aelinnel & Aelinnel & Tide-path key; host-right bough; salt license \\
Ubral Passes & Ubral & Aeler/Ubral & Pass-ring; shelter writ; guide's braid \\
Thing-Holm (Linn) & Linn & Linn & Thing ruling; harbor-green; escort writ \\
Foedus Stone (March) & Vilikari & Vilikari/Ykrul & Foedus renewal; remount allotment; hostage treaty \\
Zakov Roadstead & Zakov & Zakov & Harbor-green chit; corsair charter; pilot token \\
Black Banner Camps & Black Banner & Black Banner/Ykrul/Vilikari & Safe-conduct; remount chit; clan truce \\
Pont-du-Tithe (Belworth) & Vhasia/Viterra & Vhasia/Viterra & Dual charter; safe-conduct; duel sanction \\
Tarlington Falls & Viterra & Viterra & Lift priority; portage pass; County Thing ruling \\
Theona Moot & Theona & Theona & Moot token; Green Host immunity; ferry right \\
\hline
\end{tabular}
\caption{Gateway Control Points}
\label{tab:gateways}
\end{table}

\section{Special Rules \& Taboos}

\subsection{Mode Rules}

\subsubsection{Sea Legs (Amaranthine/Dolmis/Aberderrin)}
If Theona or Valewood 9s show up anywhere in the seed, add an omission or taboo to the leg (a missing step, an unsaid name).

\subsubsection{Passes \& Underways (Aeler)}
Any A may convert a surface route to an under-route. Diamond from Aeler always codifies outcomes---no roll---for sealed doors, breath time, and escorted segments.

\subsubsection{Rivers}
Bridges, booms, and law in Ecktoria/Viterra; reed-mazes and bell-lines in Mistlands/Linn waters.

\subsubsection{Frontier Blends}
When origin and destination disagree on law, draw two Diamonds (one from each law) and choose which you will be judged by at the end of the leg.

\subsection{Theona / Valewood 9s}

Whenever a 9 appears in a seed involving Theona or Valewood, add an omission: a missing step, an unsaid name, an unseen guest. If the 9 is a Diamond, you may break the taboo once---someone will come to collect.

\subsection{Aeler Aces}

An A can rewire routes beneath the mountains. You may swap your Spade to Aeler (keeping the other seed cards) and continue under-vault procedures.

\subsection{Echoing Omens}

Any A adds a motif you can echo later (ash on the wind, a bell heard underground, a landmark that returns elsewhere). Make the world feel like it remembers.

\section{Zakov Overlay: Gatekeepers}

Draw \textbf{one} micro-card each time you \emph{enter}, \emph{shift wharfs}, or \emph{depart} Zakov. Combine with the normal seed.

\begin{tcolorbox}[enhanced,breakable,sharp corners,boxrule=.6pt,title={\textbf{Zakov Gatekeepers Micro-Cards (draw 1)}}]
\begin{enumerate}[label=\arabic*.]
  \item \textbf{Bureaucrat Quirk} --- stamps only at \emph{eighth bell}; night filings need a \emph{Night Magistrate} countersign.
  \item \textbf{Required Stamp} --- \emph{Harbor-Green} chit for boom passage this tide; without it, DV +1 on all port checks.
  \item \textbf{Preferred Gratuity} --- figs or lamp oil; wrong gift offends, marking 1 \emph{Exposure} on your port reputation.
  \item \textbf{Audit Trap} --- ``contraband tallies''; if carrying \emph{bell-metal} or \emph{dyestuffs}, start a 4-seg \emph{Inspection} timer.
  \item \textbf{Local Courtesy} --- speak the \emph{Dock Oath} (short pledge not to knife on planks) or suffer \emph{Desperate} position in brawls.
  \item \textbf{Shortcut Rumor} --- \emph{Smuggler's Ladder} on the sea wall; spend a \diamondsuit{} (\emph{Ladder Map}) to skip one queue.
  \item \textbf{Pilot's Privilege} --- a \emph{Pilot Token} bumps you ahead on boom lift; otherwise pay tow or lose 1 segment.
  \item \textbf{Silent Court} --- the \emph{Black Bishop} is sitting tonight; legal Clubs escalate by +1 SB if you contest a ruling.
\end{enumerate}
\end{tcolorbox}

\section{Pace, Timers, and Consequences}

\begin{itemize}
\item Road/river leg: 6-segment timer; +2 segments if crossing a law boundary (new passports, new prayers).
\item Sea strait/mountain pass: 8-segment timer; --2 segments if you spend a Diamond that codifies priority (pilotage, pass, convoy).
\item Shadow corridor: 6-segment timer; on any A, add a free omen to echo later.
\item Failure defaults: delay (lose time; an enemy advances), diversion (alternate route; new Club), or debt (promise, fine, or favor owed to the authority that controls the gate).
\end{itemize}

\noindent\textbf{SB Note.} Re-rolling 1s does \emph{not} erase their SB; any new 1s on the re-roll add more SB.

\section{Story Beat (SB) Menus}

\subsection{Universal SB Spend Menu}
\begin{itemize}
  \item \textbf{1 SB} --- Shave time/position: add +1 segment to the current timer \emph{or} drop Position one step \emph{or} impose a small fee/toll \emph{or} mark 1 step of \emph{Asset Exposure}.
  \item \textbf{2 SB} --- Escalate the pressure: upgrade the active \clubsuit{} (e.g., ``inspection'' $\rightarrow$ ``seizure attempt'') \emph{or} force a \emph{choice of Diamonds} (pick which law you'll be judged by) \emph{or} tick a relevant \emph{Front} by 1.
  \item \textbf{3 SB} --- Lock a gate/taboo: close a route (boom drops; pass closes) until a named \diamondsuit{} is spent \emph{or} invoke a regional taboo (Theona/Valewood 9s; oath backlash) with teeth \emph{or} auto-trigger a \emph{Debt Flip} on a Diamond already in play.
  \item \textbf{4+ SB} --- Reshape the leg: rewrite the route (Aeler under-way reroute; sea lane reversal), start a \emph{new} 4--6 segment complication timer \emph{and} tick a \emph{Front} by +1; demand immediate payment on a Patron/Asset Exposure (lose it if refused).
\end{itemize}
\textit{End-of-leg conversion (optional):} Unspent SB convert to \emph{world motion} at a rate of \textbf{3 SB $\rightarrow$ tick 1 Front} (GM choice).

\subsection{Sea (Amaranthine/Dolmis/Aberderrin)}
\begin{itemize}
  \item \textbf{1 SB:} Fogfall nips visibility; DV +1 for the next pilotage check.
  \item \textbf{2 SB:} Boom shifts; pilot required (spend a \emph{pilotage} \diamondsuit{} or delay +2 segments).
  \item \textbf{3 SB:} Lantern-law change mid-leg; choose: pay fine \emph{or} reroute timer +2 \emph{or} spend \emph{Convoy Letter}.
  \item \textbf{4 SB:} Squall wall: split the convoy; lose an escort unless \emph{Moot token} or \emph{Escort} Diamond is spent.
\end{itemize}

\subsection{Passes \& Underways (Aeler)}
\begin{itemize}
  \item \textbf{1 SB:} White wind; breath-time check (mark Fatigue or slow +1 segment).
  \item \textbf{2 SB:} Engineer requisition; surrender cargo slot \emph{or} coin \emph{or} DV +1 until the gate.
  \item \textbf{3 SB:} Under-audit; must go below (swap Spade to Aeler) unless you spend \emph{Key-Writ}.
  \item \textbf{4 SB:} Cave-in detour adds a fresh 4-segment timer unless \emph{Underway Pass} is spent.
\end{itemize}

\subsection{Rivers (Belworth \& friends)}
\begin{itemize}
  \item \textbf{1 SB:} Sandbar drift; portage a beat or pay for a tow.
  \item \textbf{2 SB:} Lift strike at Tarlington; need \emph{Bridge/Portage Pass} or delay +2 segments.
  \item \textbf{3 SB:} Cross-mouth embargo (Marcott/Fairport); choose which harbor's law applies (draw/commit its \diamondsuit).
  \item \textbf{4 SB:} Ice-break or flood surge; split cargo (random item at risk) unless an \emph{Escort} Diamond is spent.
\end{itemize}

\subsection{Shadow Corridors (Valewood/Mistlands)}
\begin{itemize}
  \item \textbf{1 SB:} Bell-line stutter; next guidance roll at Disadvantage (or DV +1).
  \item \textbf{2 SB:} Omission manifests; something necessary is ``missing'' until \emph{Ward-salt} or \emph{Way-cord} is spent.
  \item \textbf{3 SB:} Host-law claims you as \emph{host/guest}; accept a service timer or pay with a memory (lose a minor Contact until end of leg).
  \item \textbf{4 SB:} Wraith procession crosses your path; create a 6-segment \emph{Don't Look Back} timer; spending \emph{Lantern Writ} downgrades it to 4.
\end{itemize}

\subsection{Steppe Frontiers (Ykrul/Vilikari/Black Banner)}
\begin{itemize}
  \item \textbf{1 SB:} Remount fatigue; lose 1 ``fresh mount'' use or slow +1 segment.
  \item \textbf{2 SB:} Hostage protocol; must place/accept a token-kin until the next gate (or spend \emph{Safe-hostage}).
  \item \textbf{3 SB:} Foedus recall; your papers conflict---pick Ykrul \emph{Paiza} or Vilikari \emph{Foedus Seal} and offend the other.
  \item \textbf{4 SB:} Muster checkpoint; disarm unless traveling under \emph{Standard Protection} or \emph{Khagan's Writ}.
  \item \textbf{Black Banner Specific:} Clan feud erupts; choose sides or pay double tolls.
\end{itemize}

\subsection{Ports \& Forums (Marcott, Fairport, Silkstrand, Thepyrgos, Zakov)}
\begin{itemize}
  \item \textbf{1 SB:} Clerk error; pay a ``stamp'' fee or answer questions (DV +1).
  \item \textbf{2 SB:} Rival petition; they jump the queue unless you spend a Warrant/Letters \diamondsuit{}.
  \item \textbf{3 SB:} Audit cascade; open a new 4-segment \emph{Inspection} timer tied to cargo/ledger.
  \item \textbf{4 SB:} Edict flare; convert today's \clubsuit{} into a temporary ``law'' everyone must obey or take on a \emph{Debt Flip}.
  \item \textbf{Zakov Specific:} Silent Court ruling; legal complications escalate by +1 SB.
\end{itemize}

\subsection{Diamond \& Exposure Interactions}
\begin{itemize}
  \item \textbf{2 SB:} Flip any \diamondsuit{} spent this leg: its \emph{Debt Flip} is collected now (not later).
  \item \textbf{2 SB:} Mark 1 step on a named Asset/Patron's \emph{Exposure} track.
  \item \textbf{3 SB:} Downgrade an \emph{Intricate} success to \emph{Detailed} (remove the flourish or DV −1 benefit).
  \item \textbf{4 SB:} Seize or suspend a visible Diamond's benefit until another is played (e.g., convoy letter overridden by emergency edict).
\end{itemize}

\section{Worked Itineraries}

\subsection{West-to-East Coastal Haul (Kahfagia $\rightarrow$ Viterra)}

\textbf{Leg 1: Kassamira $\rightarrow$ Ecktoria}

Seed: Ecktoria Spade (Aqueduct arcades), Ecktoria Heart (Coin-house factor), Wilds Club (gale), Kahfagia Diamond (convoy letter).

Clock: 6. Read: cross under coin-house eyes with convoy priority; bad weather strains schedules.

\textbf{Leg 2: Ecktoria $\rightarrow$ Silkstrand}

Seed: Acasia Spade (Three-Queens Bridge), Acasia Heart (Dyers' Guildmistress), Acasia Club (loom strike), Acasia Diamond (Exchange pass).

Clock: 6--8 depending on unrest.

\textbf{Leg 3: Silkstrand $\rightarrow$ Marcott}

Seed: Vhasia Spade (Pont-du-Tithe), Vhasia Heart (Parlement clerk), Vhasia Club (coin rumor), Vhasia Diamond (letters patent).

\textbf{Leg 4: Marcott $\rightarrow$ Fairport}

Seed: Viterra Spade (Fairport tideworks), Viterra Heart (shipwright), Linn Club (boom lifts), Viterra Diamond (customs seal).

Clock: 6.

\subsection{Acasia $\rightarrow$ Mistlands (Forgotten Pass + Under-Gate)}

\textbf{Leg 1: Silkstrand $\rightarrow$ Aeler Gate}

Seed: Aeler Spade (Avalanche gallery), Aeler Heart (Geometer), Aeler Club (Engineer requisition), Aeler Diamond (Underway Pass).

\textbf{Leg 2: Gate $\rightarrow$ Mistlands}

Seed: Mistlands Spade (Bell-Line levee), Mistlands Heart (Bell-warden), Mistlands Club (wraith crossing), Mistlands Diamond (Ward-salt).

\subsection{Thin Shore $\rightarrow$ Zakov Corridor}

\textbf{Leg 1: Payden's Port $\rightarrow$ Thin Shore}

Seed: Valewood Spade (Green lane), Mistlands Heart (Protectorate clerk), Mistlands Club (bell failure), Mistlands Diamond (Lantern writ).

\textbf{Leg 2: Thin Shore transit}

Seed: Valewood Spade (Unfound stile), Valewood Heart (Path-warden), Valewood Club (Sweet wind), Valewood Diamond (Way-cord).

Rule of 9s applies.

\subsection{Frontier Mediation (Ykrul $\leftrightarrow$ Vilikari)}

\textbf{Leg: Foedus Stone parley}

Seed: Vilikari Spade (Foedus Stone), Ykrul Heart (Noyan envoy), Wilds Club (rasputitsa), Vilikari Diamond (Foedus Renewal).

Blend: also draw a Ykrul Diamond (Paiza). Choose which law saves your schedule.

\subsection{Worked Itinerary: Thin Shore $\rightarrow$ Zakov $\rightarrow$ Theona (Corsair Jobs)}

\textbf{Leg 1: Payden's Port $\rightarrow$ Thin Shore (Shadow Corridor)}
\begin{quote}
Seed: Valewood \spadesuit{} (Green lane / Unfound stile); Mistlands \heartsuit{} (Protectorate clerk); Mistlands \clubsuit{} (bell-line failure); Mistlands \diamondsuit{} (Lantern Writ).
\end{quote}
\textbf{Clock:} 6. \textbf{Rule of 9s:} any 9 adds an omission (missing step/name).

\textbf{Leg 2: Thin Shore Transit (Skirting toward Zakov)}
\begin{quote}
Seed: Valewood \spadesuit{} (Sea-mist arcade); Valewood \heartsuit{} (Path-warden); Valewood \clubsuit{} (Sweet wind); Valewood \diamondsuit{} (Way-cord).
\end{quote}
\textbf{Clock:} 6. \textbf{Note:} spending \emph{Way-cord} negates one \emph{Sweet wind} lie.

\textbf{Leg 3: Approach to Zakov (Roadstead \& Booms)}
\begin{quote}
Seed: Zakov \spadesuit{} (Boomhouse or Red Wharf); Zakov \heartsuit{} (Pilot-Matron \emph{or} Night Magistrate); Zakov \clubsuit{} (Boom Drop \emph{or} Customs Sweep); Zakov \diamondsuit{} (Harbor-Green Chit \emph{or} Pilot Token).
\end{quote}
\textbf{Clock:} 6--8 (apply the Gatekeepers overlay draw on arrival). \textbf{If a 9 shows:} \emph{Missing Ninth}---remove one critical step/name until paid in favors.

\textbf{Leg 4: Corsair Job Inside Zakov (Pick 1)}
\begin{quote}
A) \textbf{Lift a Hull from Drydock Four}\\
Zakov \spadesuit{} (Drydock Four); Zakov \heartsuit{} (Corsair Quartermaster); Zakov \clubsuit{} (Admiralty Audit); Zakov \diamondsuit{} (Shipwright's Lien Release).\\
\textbf{Goal:} tow a seized hull at night; \emph{Lien Release} avoids bond but flips a debt later.
\end{quote}

\begin{quote}
B) \textbf{Court the Black Bishop for Indulgence}\\
Zakov \spadesuit{} (Black Bell Tower); Zakov \heartsuit{} (Black Bishop); Zakov \clubsuit{} (Bounty Proclamation); Zakov \diamondsuit{} (Magistrate's Hush).\\
\textbf{Goal:} seal a case/docket so a rival can't use it. \textbf{Hook:} rumor must be paid (coin or gossip).
\end{quote}

\begin{quote}
C) \textbf{Smuggler's Ladder Run}\\
Zakov \spadesuit{} (Lantern Ladder); Zakov \heartsuit{} (Lampman); Zakov \clubsuit{} (Informant Flip); Zakov \diamondsuit{} (Smuggler's Ladder Map).\\
\textbf{Goal:} bypass boom eyes with a wall-stair. \textbf{Complication:} your tip was sold twice.
\end{quote}

\textbf{Leg 5: Zakov $\rightarrow$ Theona (Isles \& Moot)}
\begin{quote}
Seed: Theona \spadesuit{} (Uncounted Bridge); Theona \heartsuit{} (Matron of Wells \emph{or} Moot Envoy); Linn \clubsuit{} (fogfall raids); Theona \diamondsuit{} (Moot Token).
\end{quote}
\textbf{Clock:} 6--8. \textbf{Taboo:} do \emph{not} count the steps aloud; breaking it imposes \emph{Controlled} on all isle bargains this scene.

\textbf{Leg 6: Theona Contract (Pick 1)}
\begin{quote}
A) \textbf{Raid-Truce at the Skerries}\\
Theona \spadesuit{} (Tide caves); Theona \heartsuit{} (Isle Moot Envoy); Theona \clubsuit{} (Muster drum); Theona \diamondsuit{} (Raid-truce Ribbon).\\
\textbf{Goal:} secure safe passage window; failure triggers a Linn muster timer.
\end{quote}

\begin{quote}
B) \textbf{Deliver the Ledger Shard}\\
Theona \spadesuit{} (Well-yard); Theona \heartsuit{} (Matron of Wells); Zakov \clubsuit{} (Debt Call follows you); Theona \diamondsuit{} (Sanctuary Night).\\
\textbf{Goal:} hide a Zakov debt long enough to copy data; sanctuary buys time.
\end{quote}

\subsection{Steppe Passage: Black Banner Territory}

\textbf{Leg: Foedus Stone $\rightarrow$ Black Banner Camp}
\begin{quote}
Seed: Ykrul/Vilikari \spadesuit{} (Wolf Road milepost or Foedus Stone); Black Banner \heartsuit{} (Clan Elder or War Captain); Wilds \clubsuit{} (Rasputitsa or Remount Sickness); Black Banner \diamondsuit{} (Safe-conduct or Remount Chit).
\end{quote}
\textbf{Clock:} 6--8. \textbf{Complication:} Foedus recall may invalidate your papers.

\textbf{Leg: Black Banner Camp $\rightarrow$ Ykrul Territory}
\begin{quote}
Seed: Ykrul \spadesuit{} (Winter camp ring or Khagan's way-station); Ykrul \heartsuit{} (Khatun of the Ring or Noyan envoy); Frontier \clubsuit{} (Hostage protocol or Feud spark); Ykrul \diamondsuit{} (Paiza tablet or Foedus seal).
\end{quote}
\textbf{Clock:} 6--8. \textbf{Note:} Choose which law applies---Black Banner customs or Ykrul traditions.

\section{Quick Reference: Routing Options}

\subsection{From Acasia/Silkstrand}
\begin{itemize}
\item West to Ecktoria: Amaranthine Coastway (Leg 2)
\item East to Vhasia/Viterra: Amaranthine Coastway (Legs 3-4)
\item North to Ykrul/Vilikari: Steppe Frontiers (Leg 2)
\item South to Thepyrgos: River Roads (Acasian Rivers $\rightarrow$ Ecktoria $\rightarrow$ Viterra)
\end{itemize}

\subsection{From Ecktoria}
\begin{itemize}
\item West to Kahfagia: Amaranthine Coastway (Leg 1)
\item East to Acasia: Amaranthine Coastway (Leg 2)
\item North to Aeler/Mistlands: Aelerian Passes (Leg 2)
\item East to Viterra/Vhasia: Amaranthine Coastway (Legs 3-4)
\end{itemize}

\subsection{From Vhasia (West Balkans)}
\begin{itemize}
\item North to Aeler: Aelerian Passes (Leg 1)
\item South to Thepyrgos: River Roads (Leg 4)
\item West to Acasia: Amaranthine Coastway (Leg 3)
\item East to Viterra: River Roads (Belworth crossing)
\end{itemize}

\subsection{From Viterra (East Balkans, straddles seas)}
\begin{itemize}
\item North to Ubral/Aeler: Aelerian Passes (Leg 2) / Ubral Routes
\item South to Thepyrgos: River Roads (Black River)
\item West to Vhasia/Acasia: Amaranthine Coastway (Legs 3-4)
\item East to Dolmis: Dolmis Circuit (Leg 1)
\end{itemize}

\subsection{From Thepyrgos (Southwest corner)}
\begin{itemize}
\item North to Viterra: River Roads (Black River)
\item West to Amaranthine: Astroegro Straits (Leg 2)
\item East to Dolmis: Astroegro Straits (Leg 3)
\item Northwest to Vhasia: River Roads (via Viterra)
\end{itemize}

\subsection{From Ykrul/Vilikari}
\begin{itemize}
\item South to Acasia/Ecktoria: Steppe Frontiers (Leg 2)
\item West to Mistlands/Aeler: Aelerian Passes (Leg 6) or Direwood crossing
\item East to Valewood: Thin Shore corridor
\end{itemize}

\subsection{From Mistlands/Aeler}
\begin{itemize}
\item North to Ykrul: Aelerian Passes (Leg 4) or Direwood crossing
\item South to Ecktoria: Aelerian Passes (Leg 2)
\item West to Linn/Valewood: Shadow Corridors (Leg 1)
\end{itemize}

\subsection{From Ubral}
\begin{itemize}
\item South to Viterra: Ubral Routes (Leg 1)
\item North to Aeler: Ubral Routes (Leg 2)
\item East to Dolmis: Ubral Routes (Leg 3)
\end{itemize}

\subsection{From Linns}
\begin{itemize}
\item South via portage: Linn Raid Routes $\rightarrow$ Thin Coast $\rightarrow$ Dolmis
\item South via sea: Linns $\rightarrow$ Aberderrin $\rightarrow$ Mistlands/Valewood
\end{itemize}

\subsection{From Valewood}
\begin{itemize}
\item West to Mistlands: Shadow Corridors (Leg 2)
\item South to Aelinnel/Aelaerem: Green Gate Hops (Leg 3)
\item North to Thin Coast/Zakov: Shadow Corridors (Leg 1)
\end{itemize}

\end{document}
